### 2018-07-09 soc_squeezing/soc_squeezing_notes.tex
add soc-induced squeezing notes

\documentclass[aps,notitlepage,nofootinbib,11pt]{revtex4-1}

% linking references
\usepackage{hyperref}
\hypersetup{
  breaklinks=true,
  colorlinks=true,
  linkcolor=blue,
  filecolor=magenta,
  urlcolor=cyan,
}

%%% symbols, notations, etc.
\usepackage{physics,braket,bm,commath,amssymb} % physics and math
\renewcommand{\t}{\text} % text in math mode
\newcommand{\f}[2]{\dfrac{#1}{#2}} % shorthand for fractions
\newcommand{\p}[1]{\left(#1\right)} % parenthesis
\renewcommand{\sp}[1]{\left[#1\right]} % square parenthesis
\renewcommand{\set}[1]{\left\{#1\right\}} % curly parenthesis
\renewcommand{\v}{\bm} % bold vectors
\renewcommand{\c}{\cdot} % inner product
\newcommand{\bk}{\braket} % shorthand for braket notation
\newcommand{\Bk}{\Braket} % shorthand for braket notation

\renewcommand{\d}{\text{d}} % "d" for integration measure
\newcommand{\g}{\text{g}} % ground / excited electronic states
\newcommand{\e}{\text{e}}

\newcommand{\B}{\mathcal{B}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\G}{\mathcal{G}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\J}{\mathcal{J}}
\renewcommand{\L}{\mathcal{L}}
\renewcommand{\O}{\mathcal{O}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\Q}{\mathcal{Q}}

\usepackage{dsfont} % for identity operator
\newcommand{\1}{\hat{\mathds{1}}}

\usepackage[inline]{enumitem} % for inline enumeration

%%% figures
\usepackage{graphicx} % for figures
\usepackage{grffile} % help latex properly identify figure extensions
\usepackage[caption=false]{subfig} % subfigures (via \subfloat[]{})
\graphicspath{{./plots/}} % set path for all figures

% for strikeout text
% normalem included to prevent underlining titles in the bibliography
\usepackage[normalem]{ulem}

% for leaving notes in the text
\newcommand{\note}[1]{\textcolor{red}{#1}}

\begin{document}

\title{Spin-orbit-coupling-induced squeezing in the optical lattice
  clock}

\author{Michael A. Perlin}

\maketitle

These are some notes about trying to induce squeezing in an optical
lattice clock by combining spin-orbit coupling and interactions.  We
assume, for now, that
\begin{enumerate*}[label=(\roman*)]
\item atoms are fermionic in nature with an alkali-earth(-like)
  electronic structure,
\item atoms are nuclear-spin-polarized, and
\item the optical lattice is quasi-one-dimensional, with tight
  (i.e.~ground-state) confinement in transverse directions.
\end{enumerate*}

\section{Spin-orbit coupling}

We start with two-level atoms loaded into the ground band of an
optical lattice, and consider interrogating these atoms by a linearly
polarized plane-wave laser tuned to the relevant electronic (atomic)
transition, with wavenumber projection $\phi$ onto the lattice axis.
After performing a gauge transformation which shifts the on-axis
momenta of ground (excited) states by $\phi/2$ ($-\phi/2$), and
eliminating (i.e.~turning off) the interrogation laser, one can arrive
at the Hamiltonian
\begin{align}
  H_{\t{SOC}}
  = -2J_0\sum_{q,s} \cos\p{q+s\phi/2} \hat c_{qs}^\dag \hat c_{qs}
  \label{eq:H_SOC_start}
\end{align}
where $q$ indexes quasi-momentum along the lattice axis,
$s\in\set{\g\leftrightarrow-1,\e\leftrightarrow1}$ labels the
electronic state, $J_0$ is the (positive) ground-band tunneling rate,
$c_{q\sigma}$ is a fermionic annihilation operator, and we work in
units with the lattice spacing $a=1$.  Defining
\begin{align}
  \1_q \equiv \hat c_{q,\g}^\dag \hat c_{q,\g}
  + \hat c_{q,\e}^\dag \hat c_{q,\e},
  &&
  \sigma_q^j \equiv \sum_{\alpha,\beta}
  \hat c_{q\alpha}^\dag \sigma^j_{\alpha\beta} \hat c_{q\beta},
  \label{eq:pseudospin}
\end{align}
for Pauli matrices $\sigma^j$ (i.e.~such that $\sigma_q^j$ is a
pseudo-spin-1/2 Pauli operator on the electronic degrees of freedom)
and expanding the cosine in \eqref{eq:H_SOC_start}, we thus find
\begin{align}
  H_{\t{SOC}}
  = -\sum_q \p{\epsilon_q \1_q + \f12 h_q \sigma_q^z},
  &&
  \epsilon_q \equiv 2J_0 \cos\p{\phi/2} \cos q,
  &&
  h_q \equiv -4J_0 \sin\p{\phi/2} \sin q.
\end{align}
When all dynamics conserve the occupations of all quasi-momenta, the
first term in this Hamiltonian merely contributes a constant energy
shift which we can safely neglect, leaving us with the effective
inhomogeneous-field Hamiltonian
\begin{align}
  H_{\t{SOC}} = -\sum_q h_q s_q^z,
  &&
  s_q^z \equiv \f12 \sigma_q^z
  \label{eq:H_SOC}
\end{align}

\section{Interactions}

At ultracold temperatures inter-atomic interactions are dominated by
$s$-wave collisions, which are captured by the Hamiltonian
\begin{align}
  H_{\t{int}} = G \int \d^3x~
  \hat\psi_\e^\dag\p{x} \hat\psi_\g^\dag\p{x}
  \hat\psi_\g\p{x} \hat\psi_\e\p{x},
  &&
  G \equiv \f{4\pi a_{\e\g^-}}{m_A},
\end{align}
where $\hat\psi_\sigma$ is a fermionic field operator for atoms in
electronic state $\sigma\in\set{\g,\e}$, $a_{eg}^-$ is an scattering
length, and $m_A$ is the mass of a single atom.  Considering only
occupation of the ground band in a lattice, we can expand the field
operators as
\begin{align}
  \hat\psi_\sigma\p{x} = \sum_q \phi_q\p{x} \hat c_{q\sigma},
\end{align}
for momentum-space (Bloch) wavefunctions $\phi_q$ and fermionic
annihilation operators $\hat c_{q\sigma}$.  The interaction
Hamiltonian then becomes
\begin{align}
  H_{\t{int}} = G \sum_{p,q,r,s} K^{pq}_{rs}
  \hat c_{r,\e}^\dag \hat c_{s,\g}^\dag \hat c_{q,\g} \hat c_{p,\e},
  &&
  K^{pq}_{rs} \equiv \int \d^3x~
  \phi_r\p{x}^* \phi_s\p{x}^* \phi_q\p{x} \phi_p\p{x}.
\end{align}
In a lattice with $L$ sites centered on positions $x_j$ (i.e.~for
$j=1,\cdots L$ indexing lattice site) and localized (Wannier)
wavefunctions $w_j$, we can expand
$\phi_p\p{x}=L^{-1/2}\sum_je^{-ipx_j}w_j\p{x}$, which implies
\begin{align}
  K^{pq}_{rs} = L^{-2} \sum_{j,k,\ell,m} \int \d^3x~
  \exp\sp{-i\p{px_j+qx_k-rx_\ell-sx_m}}
  w_\ell\p{x}^* w_m\p{x}^* w_k\p{x} w_j\p{x}.
\end{align}
If we now assume that we can neglect integrals without $j=k=\ell=m$,
which is equivalent to neglecting inter-site interactions and
interaction-assisted hopping, then we have
\begin{align}
  K^{pq}_{rs} \approx L^{-2} \sum_j \int \d^3x~
  \exp\sp{-i\p{p+q-r-s}x_j} \abs{w_j\p{x}}^4,
\end{align}
where a stationary phase approximation now forces $p+q-r-s=0$ (mod
$2\pi$ in units with the lattice spacing $a=1$), which is equivalent
to conservation of total momentum.  In terms of the Kronecker delta
$\delta_{p+q,r+s}$ enforcing $p+q=r+s$, it follows that
\begin{align}
  K^{pq}_{rs} \approx \delta_{p+q,r+s} L^{-2}
  \sum_j \int \d^3x~ \abs{w_j\p{x}}^4
  = \delta_{p+q,r+s} L^{-1} \int \d^3x~ \abs{w_0\p{x}}^4,
\end{align}
which motivates the definitions
\begin{align}
  U \equiv G \int \d^3x~ \abs{w_0\p{x}}^4,
  &&
  \xi \equiv \f{U}{L},
\end{align}
in order to express the interaction Hamiltonian in the simple form
\begin{align}
  H_{\t{int}} = \xi \sum_{p,q,r,s}
  \delta_{p+q,r+s} \hat c_{r,\e}^\dag \hat c_{s,\g}^\dag
  \hat c_{q,\g} \hat c_{p,\e}.
  \label{eq:H_int_full}
\end{align}

\section{Collective spin model}

If the single-particle energy spacings are much greater than the
strength of inter-particle interactions, or equivalently if
$J_0\gg U$, then by the secular approximation we can neglect terms in
$H_{\t{int}}$ which do not conserve the sum of single-particle
energies.  By solving this single-particle energy conservation
condition, one finds that we must have $\set{p,q}\ne\set{r,s}$; that
is, atoms in different clock states can only interact via
\begin{enumerate*}[label=(\roman*)]
\item direct density-density terms, and
\item terms which exchange their momenta.
\end{enumerate*}
The surviving terms in \eqref{eq:H_int_full} are thus
\begin{align}
  H_{\t{int}}
  = \xi \sum_{p,q}
  \p{\hat c_{p,\e}^\dag \hat c_{q,\g}^\dag \hat c_{q,\g} \hat c_{p,\e}
    + \hat c_{q,\e}^\dag \hat c_{p,\g}^\dag \hat c_{q,\g} \hat c_{p,\e}}
  = \xi \sum_{p,q}
  \p{\hat c_{q,\g}^\dag \hat c_{q,\g} \hat c_{p,\e}^\dag \hat c_{p,\e}
    - \hat c_{q,\e}^\dag \hat c_{q,\g} \hat c_{p,\g}^\dag \hat c_{p,\e}},
\end{align}
where we can use the single-particle pseudo-spin-1/2 Pauli operators
in \eqref{eq:pseudospin} to write
\begin{align}
  H_{\t{int}} = \xi \sum_{p,q}
  \sp{\p{\f{\1_q-\sigma_q^z}{2}} \p{\f{\1_p+\sigma_p^z}{2}}
    - \sigma_q^+ \sigma_p^-}.
\end{align}
Up to a global shift in energy, this result is equivalently
\begin{align}
  H_{\t{int}}
  = - \f14 \xi \sum_{p,q} \v{\sigma}_q\c\v{\sigma}_p
  = - \xi \sum_{p,q} \v s_q \c\v s_p
  = - \xi \v S \c \v S,
\end{align}
where
\begin{align}
  \v{\sigma}_p
  \equiv \p{\sigma_p^x,\sigma_p^y,\sigma_p^z},
  &&
  \v s_p \equiv \p{s_p^x,s_P^y,s_p^z},
  &&
  \v S \equiv \f12 \sum_p \v{\sigma}_p.
  \label{eq:H_int}
\end{align}
For a fixed particle number $N$, the collective spin Hamiltonian
$H_{\t{int}}$ has electronic eigenstates $\set{\ket{Sm}}$ with total
(pseudo-)spin $S$ and projection $m$ onto a quantization axis; the
corresponding energies are $\bk{Sm|H_{\t{int}}|S,m}=-\xi S\p{S+1}$.
The ground-state $S=N/2$ manifold is spanned by the Dicke states
\begin{align}
  \ket{m} \equiv \ket{N/2,m}
  = {N \choose N/2+m}^{-1/2} S_+^{N/2+m} \ket{\g}^{\otimes N},
  &&
  S_+ \equiv \sum_n \sigma_n^+,
\end{align}
for which we note that $N/2+m=0,1,\cdots,N$ is the number of
electronic state excitations in the state $\ket{m}$.  Above this
ground-state manifold with an energy gap of $N\xi$ lies the manifold
of $S=N/2-1$ states, which is spanned by the orthonormal spin-wave
states
\begin{align}
  \ket{mk}
  \equiv \ket{N/2-1,m,k}
  \equiv {N-1 \choose \p{N/2-m}\p{N/2-m+1}}^{-1/2}
  \sum_{n=1}^N e^{2\pi ikn/N} s_n^+ \ket{N/2,m-1},
\end{align}
where $k=1,2,\cdots,N-1$; $n$ indices an individual atom; and
$s_n^+\equiv\hat c_{n,\e}\hat c_{n,\g}$
\cite{swallows2011suppression}.  \note{[We should comment about how we
  do not actually achieve $J_0\gg U$ in practice, but that we can
  nonetheless validate the spin model using numerics.]}

\section{Mean-field treatment}

To sum up our developments thus far, spin-orbit coupling gives rise to
an effective inhomogeneous field, while $s$-wave interactions give
rise to a collective spin interaction, which means that the system is
characterized by the Hamiltonian \note{[MP: I removed the mention of
  non-uniform coupling (i.e.~through $\xi_{ij}$), as I feel it is
  unnecessary given the preceding section on the collective spin
  model.  The uniform all-to-all coupling in \eqref{eq:H_int} is exact
  under the assumptions of the spin model, so there is no need to
  ``re-derive'' it by throwing out inhomogeneous couplings and taking
  $\xi_{ij}\to\xi\equiv\overline{\xi_{ij}}$.]}
\begin{align}
  H = H_{\t{int}} + H_{\t{SOC}} = -\xi\v S\c\v S - \sum_q h_q s_q^z,
  \label{eq:H_full}
\end{align}

\note{[MP] Note: ``eqnarray'' environment is generally not recommended
  for typesetting math, as it has inconsistent spacing in addition to
  some other problems (discussed a bit
  \href{http://www.tug.org/pracjourn/2006-4/madsen/madsen.pdf}{here},
  if you're interested).  The ``align'' environment, on the other
  hand, is provided by math packages for the purposes of typesetting
  math, lacks spacing problems, and generally has more functionality.}

This collective Heisenberg model is a celebrated model for magnetism,
and is closely related to the Richardson-Gaudin models for
superconductivity based on the Anderson pseudo-spin mapping
\note{[cite?]}.  So before demonstrating how to generate an effective
one-axis squeezing Hamiltonian based on Eqn.~\eqref{eq:H_full}, we
first try to develop better understanding of this Hamiltonian.  In
fact, this model is well-studied at the mean-field (MF) level as given
by
\begin{align}
  H_{\t{MF}} = -\sum_q \v s_q \c \v B_q,
  &&
  \v B_q \equiv \xi\v{\mathcal S} + \v h_q,
  &&
  \v{\mathcal S} \equiv \bk{\v S},
  &&
  \v h_q \equiv \p{0,0,h_q}.
  \label{eq:mean_field}
\end{align}
\note{[MP: I removed a factor of 2 in front of $\xi$, as I believe it
  was a mistake.  Was this correct?  If the factor of 2 was indeed
  incorrect, please make sure that there are no extra factors of 2 in
  Peiru's codes.]}
% MP: as an additional side note, I personally recommend the use of
% \braket{X} or (\bk{X}, which I defined as a shorthand) in place of
% \langle X \rangle; I find this is a little easier to read.
% Similarly, you can use \ket{X} (\bra{X}) in place of |X\rangle in
% place of (\langle X|), and \op{X}{Y} in place of |X\rangle\langle Y|
% (where "op" stands for "outer product")

The behavior of this model is determined by a competition between the
non-local MF coupling through $\xi$ and an effective inhomogeneous
axial field $\v h_q$ whose strength has a variance
$8J_0^2\sin^2\p{\phi/2}$.  Fig.~\ref{fig:MFcontrast} shows the
dynamics of the system as computed by numerical simulation of
\eqref{eq:mean_field}, after initialization with all spins pointing
along the $x$ axis, i.e.~in the state
$\bigotimes_q\p{\ket{q,\g}+\ket{q,\e}}/\sqrt2$. \note{[MP: I added
  text specifying the initial state; is it correct?]}  As the
interaction strength $\xi$ increases, the contrast
$\mathcal C\equiv2\abs{\mathcal S_+}/N$, \note{[MP: I propose writing
  $\mathcal C\equiv\sqrt{\mathcal S_x^2+\mathcal S_y^2}/\p{N/2}$;
  while longer, I feel like it makes the physical interpretation of
  $\mathcal C$ more clear.]} i.e.~the normalized magnitude of the
collective spin vector in the $x$-$y$ plane, gets stabilized against
dephasing induced by the effective inhomogeneous axial field $\v h_q$.
Furthermore, when $\xi$ is below a critical value $\xi_{\t{c}}$, the
contrast exhibits some oscillatory behavior, but ultimately decays at
long times.  Above the critical value $\xi_{\t{c}}$, however, the
contrast reaches a non-zero steady state, with oscillatory behavior
that is damped with increasing $\xi$.  These two phases can be
understood through the underlying integrability of
\eqref{eq:mean_field}.  Using the techniques in \note{[cite]}, the
steady state contrast $\bar{\mathcal C}$ at the MF level can be
obtained analytically as
\begin{align}
  \bar{\mathcal C} =
  \begin{cases}
    \sqrt{1-\sp{4J_0\sin\p{\phi/2}/N\xi}^2},
    & \text{if}~ \abs{4J_0\sin(\phi/2)/N\xi} < 1 \\
    0,
    & \text{otherwise}
  \end{cases}
\end{align}

\begin{figure}
\centering
\includegraphics[width=3.5in]{MFcontrast.pdf}
\caption{
The contrast as a function of time and interaction strength.
}
\label{fig:MFcontrast}
\end{figure}

Comparing the MF simulation results with the quantum simulation
results \note{[MP: In my experience, ``quantum simulation'' refers to
  that which is performed with a quantum simulator, i.e.~via
  experiment.  I assume that's not what you meant here, and that these
  are actually some sort of Monte Carlo / quantum trajectory / TWA
  simulations.  Can we specify the method?]}, however, we find that
the contrast decreases rather than remain as a constant even if
$\xi\gg\xi_{\t{c}}$, as shown in Fig.~\ref{fig:EffectiveSqueezing}.

\begin{figure}
\centering
\includegraphics[width=3in]{Fig33.pdf}
\caption{Contrast comparison among the MF simulation results, quantum simulation results and analytic expressions:
(a) Good agreement between MF simulation (solid line) and quantum simulation (dashed line) at short time,
while  the discrepancy occurs at longer time due to the quantum fluctuations.
(b-c) Benchmark the approximated one-axis twisting analytical expressions (dashed line) with the quantum treatment (solid line).
}
\label{fig:EffectiveSqueezing}
\end{figure}

\note{[MP: While it may be nice for understanding some properties of
  the Hamiltonian in \eqref{eq:H_full}, I do not think that this text
  about contrast and mean-field treatment adds much to our goal of
  spin squeezing.  Do we need to include this material here?  I would
  cut this section out for now.]}

\section{Effective spin squeezing}
\label{sec:OAT}

In order to demonstrate spin squeezing behavior induced by spin-orbit
coupling, we fix the total particle number $N$, consider initial
states with total spin $S=N/2$, and write the total system Hamiltonian
as
\begin{align}
  H_{\t{free}} = H_{\t{int}} + H_{\t{SOC}}
  = -\xi\v S\c\v S - \sum_n h_n s_n^z
  = -\xi\v S\c\v S - \bar h S_z - \sum_n b_n s_n^z,
\end{align}
where we have separated out the mean and residual effective fields
\begin{align}
  \bar h \equiv \f1N \sum_n h_n,
  &&
  b_n \equiv h_n - \bar h.
\end{align}
The subscript on $H_{\t{free}}$ refers to the fact that this
Hamiltonian corresponds to free, undriven evolution of (interacting)
atoms on the lattice.  The collective spin operators $\v S\c\v S$ and
$S_z$ both preserve total spin $S$ and its projection $m$ onto a
quantization axis, which makes the Dicke states a good basis for
treating $H_{\t{free}}$ (i.e.~within the $S=N/2$ manifold).  While the
residual effective field term $\sum_n b_n s_n^z$ generally takes
states outside the $S=N/2$ manifold, we can treat this spin-changing
process perturbatively when the residual effective fields $b_n$ are
small compared to the collective spin gap, i.e.~when
$\max_n\abs{b_n}\ll N\abs{\xi}$.  Such a perturbative treatment yields
an effective Hamiltonian of the form
$H_{\t{free,eff}} = H_{\t{free,eff}}^{(1)} + H_{\t{free,eff}}^{(2)}$
with
\begin{align}
  H_{\t{free,eff}}^{(1)}
  &= \sum_m \op{m} H \op{m}
  = -\xi \v S \c \v S - \bar h S_z, \\
  H_{\t{free,eff}}^{(2)}
  &= -\sum_{\substack{m\\S<N/2}}
  \frac{\abs{\bk{Sm|H|m}}^2}{E_{Sm}^{\t{free}}-E_m^{\t{free}}} \op{m}
  = -\sum_{\substack{m\\S<N/2}}
  \frac{\abs{\sum_n b_n \bk{Sm|s_n^z|m}}^2}
  {E_{Sm}^{\t{free}}-E_m^{\t{free}}} \op{m},
\end{align}
where $E_\psi^{\t{free}}\equiv\bk{\psi|H_{\t{free}}|\psi}$ and we have
used the fact the spin projection $m$ is conserved by $H$.  As the
individual spin operators $s_n^z$ cannot change the total spin $S$ by
more than 1, we need only consider $S=N/2-1$ spin-wave states when
evaluating the sum over $S<N/2$ in $H_{\t{free,eff}}^{(2)}$ above.
These spin-wave states couple to the Dicke states via the matrix
elements of the form (see section 2 in the supporting materials of
\cite{swallows2011suppression})
\begin{align}
  \bk{m|s_n^z|mk} = e^{2\pi i k n/N} \sqrt{\f{(N/2)^2-m^2}{N^2 (N-1)}},
\end{align}
which implies
\begin{align}
  H_{\t{free,eff}}^{(2)}
  = -\sum_{m,k}
  \frac{\abs{\sum_n b_n \bk{mk|s_n^z|m}}^2}{N\xi} \op{m}
  = -\sum_m \f{(N/2)^2-m^2}{N^3(N-1)\xi}
  \sum_k \abs{\sum_n b_n e^{2\pi ikn/N}}^2 \op{m}.
\end{align}
Using the fact that $\sum_nb_n=0$, we can expand
\begin{align}
  \sum_k \abs{\sum_n b_n e^{2\pi ikn/N}}^2
  = \sum_{k=1}^{N-1} \sum_{n,n'} b_n b_{n'} e^{2\pi ik\p{n-n'}/N}
  = \sum_{n,n'} b_n b_{n'} \sum_{k=0}^{N-1} e^{2\pi ik\p{n-n'}/N},
\end{align}
where the sum over $k$ vanishes for $n\ne n'$ and equals $N$ when
$n=n'$, so
\begin{align}
  \sum_k \abs{\sum_n b_n e^{2\pi ikn/N}}^2
  = N \sum_n b_n^2 = N^2 \tilde h^2,
  &&
  \tilde h^2 \equiv \f1N \sum_n b_n^2 = \f1N \sum_n \p{h_n - \bar h}^2.
  \label{eq:sum_knn}
\end{align}
We therefore have that
\begin{align}
  H_{\t{free,eff}}^{(2)}
  = -\sum_m \f{\p{N/2}^2-m^2}{N\p{N-1}\xi}~ \tilde h^2 \op{m},
\end{align}
where the term $\p{N/2}^2$ term contributes a global energy shift
which we can neglect, while the $m^2$ term is proportional to
$m^2\op{m}=S_z^2$.  In total, the effective Hamiltonian is thus
\begin{align}
  H_{\t{free,eff}}
  = -\xi \v S \c \v S - \bar h S_z + \chi S_z^2,
  &&
  \chi \equiv \f{\tilde h^2}{N\p{N-1}\xi}.
\end{align}
This effective Hamiltonian preserves total spin $S$, which means that
the $\v S\c\v S$ term contributes only a constant energy shift which
we can neglect.  Furthermore, at temperatures $T\gg J_0$ the mean
effective field $\bar h=0$, while for $T\sim J_0$ the collective
rotation due to the $S_z$ term can be neglected by use of a rotating
frame.  The effective Hamiltonian therefore simplifies to the one-axis
twisting Hamiltonian
\begin{align}
  H_{\t{OAT}} = \chi S_z^2.
  \label{eq:H_OAT}
\end{align}
An example set of parameters and values for one-axis twisting via
$H_{\t{OAT}}$ with strontium-87 is provided in Table
\ref{tab:parameters}, along with the optimal one-axis twisting
squeezing time
$t_{\t{opt}}^{\t{OAT}}\approx1.1N^{-2/3}\chi^{-1}\sim\eta
N^{1/3}=N^{4/3}/L$.  Lattice depths are provided in units of the
lattice recoil energy $E_R$.

\begin{table}[h]
  \centering
  \caption{Example parameters for spin squeezing via one-axis
    twisting.}
  \label{tab:parameters}
  \begin{tabular}{|l|c|l|}
    \hline
    Parameter & Symbol & Value \\ \hline\hline
    Primary lattice depth & $V_0$ & 3 $E_R$ \\
    Transverse lattice depths & $V_T$ & 80 $E_R$ \\
    Lattice sites & $L$ & 100 \\
    Atom number & $N$ & 50 \\
    SOC strength & $\phi$ & $\pi/50$ \\ \hline\hline
    Tunneling rate & $J_0$ & $\approx300\times2\pi$ Hz \\
    On-site interaction strength & $U$ & $\approx1.4\times2\pi$ kHz \\
    Collective interaction strength & $\xi$
    & $\approx14\times2\pi$ Hz \\
    Squeezing strength & $\chi$ & $\approx21\times2\pi$ mHz \\
    Optimal squeezing time & $t_{\t{opt}}^{\t{OAT}}$
    & $\approx0.66$ seconds \\ \hline
  \end{tabular}
\end{table}

\section{Two-axis twisting: pulsed drive}

While one-axis twisting provides a spin fluctuation noise floor which
scales with atom number as $\sim1/N^{2/3}$, it is possible to beat
this limit and achieve a noise floor scaling as $\sim1/N$ via
so-called two-axis twisting.  We can transform the one-axis twisting
Hamitlonian in \eqref{eq:H_OAT} into an effective two-axis twisting
Hamiltonian by using a pulsed drive sequence developed in
ref.~\cite{liu2011spin}.  The necessary drive sequence requires the
pulsed application of a clock laser, which turns on a collective spin
Hamiltonian
\begin{align}
  H_{\t{clock}} = -\Omega \sum_n s_n^x = -\Omega S_x.
\end{align}
If $\abs{\Omega}\gg N\chi$, then we can apply $H_{\t{clock}}$ for
short times $\theta/\abs{\Omega}\ll\p{N\chi}^{-1}$ (during which we
can neglect $H_{\t{eff}}$) to realize collective rotations of the form
$\exp\p{\pm i\theta S_x}$.  Thus acting with $\exp\sp{i\p{\pi/2}S_x}$,
waiting for a time $2\tau$, acting with $\exp\sp{-i\p{\pi/2}S_x}$, and
waiting for a time $\tau$ then realizes the unitary
\begin{align}
  U_1 = \exp\p{-i\tau\chi S_z^2} \exp\p{-i\f{\pi}{2}S_x}
  \exp\p{-i2\tau\chi S_z^2} \exp\p{i\f{\pi}{2}S_x}
  = \exp\p{-i\tau\chi S_z^2} \exp\p{-i2\tau\chi S_y^2}.
\end{align}
If $\tau\chi\ll1$, then
\begin{align}
  U_1 \approx \exp\sp{-i\tau\chi\p{S_z^2 + 2S_y^2}}
  = \exp\sp{-i\tau\chi\p{S^2 + S_y^2 - S_x^2}}
  \simeq \exp\sp{-i\tau\chi\p{S_y^2 - S_x^2}},
\end{align}
where $\simeq$ denotes equality up to an overall phase, which holds
above because $S^2$ is conserved by all dynamics.  Repeating the above
sequence $n$ times for a time $t=3n\tau$ thus results in the two-axis
twisting unitary
\begin{align}
  U = \exp\sp{-in\tau\chi\p{S_y^2 - S_x^2}}
  = \exp\sp{-i\p{t/3}\chi\p{S_y^2 - S_x^2}},
\end{align}
which is equivalent to evolving for a time $t$ under the two-axis
twisting Hamiltonian
\begin{align}
  H_{\t{TAT}}^{y,x} = \f{\chi}{3}\p{S_y^2 - S_x^2}.
  \label{eq:H_TAT_pulse}
\end{align}
Despite the factor of 1/3 in the effective two-axis twisting strength,
the optimal squeezing time for this two-axis twisting Hamiltonian is
generally close to that for one-axis twisting with $H_{\t{OAT}}$ in
\eqref{eq:H_OAT} \cite{liu2011spin}.

\note{[MP: In practice, of course, the gates $\exp\p{i\theta S_x}$
  take a nonzero amount of time to implement.  For the purposes of
  simulation, we can probably approximate these gates by square pulses
  centered on the appropriate time.  I believe that experiments are
  capable of achieving $\Omega\sim1\times2\pi~\t{kHz}$.  I can do an
  error analysis on this sequence to determine the effect of the
  $\O\p{\tau^2\chi^2}$ terms in the exponent of $U_1$, but it is
  probably simpler to just try this out in simulations.]}

\section{Two-axis twisting: continuous drive}

Rather than pulsing the clock laser, we can apply a continuous drive
to realize the Hamiltonian
\begin{align}
  H_{\t{drive}} = H_{\t{free}} + H_{\t{clock}}
  = -\xi\v S\c\v S - \bar h S_z - \Omega S_x - \sum_n b_n s_n^z,
\end{align}
where the drive strength $\Omega$ may generally depend on time.  As
shown in Appendix \ref{sec:effective_squeezing_drive}, the addition of
a drive with $\abs{\Omega}\ll N\abs{\xi}$ does not modify the
squeezing behavior of the spin-orbit coupling terms in
$H_{\t{drive}}$, whose effective action on the $S=N/2$ manifold is
captured by
\begin{align}
  H_{\t{drive,eff}} = -\bar h S_z - \Omega S_x + \chi S_z^2.
  \label{eq:H_drive_eff}
\end{align}
To follow the continuous-drive two-axis twisting prescription in
ref.~\cite{huang2015twoaxis}, we now
\begin{enumerate*}[label=(\roman*)]
\item blue-detune the clock laser by $\bar h$ and shift into an
  appropriate rotating frame to eliminate the $\sim S_z$ term in
  \eqref{eq:H_drive_eff},
\item modulate the drive as $\Omega\p{t}=\beta\omega\cos\p{\omega t}$
  with $\omega\gg N\chi$,
\item move into the rotating frame of $-\Omega\p{t}S_x$, and finally
\item shift $H_{\t{drive,eff}}$ by the constant
  $\p{\chi/2}\sp{\J_0\p{2\beta}-1} S^2$, where $\J_0\p{x}$ is the
  zero-order Bessel function of the first kind evaluated at $x$.
\end{enumerate*}
After a secular approximation which relies on $\omega\gg N\chi$, this
procedure results in the effective Hamiltonian
\begin{align}
  H_{\t{drive,eff}}^{\t{mod}}
  = \f{\chi}{2} \sp{3\J_0\p{2\beta}-1} S_z^2
  + \f{\chi}{2} \sp{1-\J_0\p{2\beta}} \p{S_z^2 - S_x^2},
\end{align}
where choosing the modulation index $\beta$ such that
$\J_0\p{2\beta}=1/3$ or $\J_0\p{2\beta}=-1/3$ respectively results in
the two-axis twisting Hamiltonians
\begin{align}
  H_{\t{TAT}}^{z,x} = \f{\chi}{3} \p{S_z^2 - S_x^2},
  &&
  H_{\t{TAT}}^{y,x} = \f{\chi}{3}\p{S_y^2 - S_x^2}.
  \label{eq:H_TAT_drive}
\end{align}
Given the set of parameters in Table \ref{tab:parameters}, the
two-axis twisting Hamiltonian $H_{\t{TAT}}^{z,x}$ can be achieved with
$\omega=20\times2\pi$ Hz and $\beta=0.9057$.  The optimal squeezing
time for this protocol is $t_{\t{opt}}^{\t{TAT}}\approx0.98$ seconds.

\section{Discussions}

\note{[MP: We should clarify in the notes what exactly the conditions
  are for the effective Hamiltonians we derive, i.e.~the spin model
  and the one-axis twisting Hamiltonian.]}

\note{[MP] Future ideas:} Make use of a cavity to enhance the one-axis
twisting strength \cite{hu2017vacuum}?

\appendix

\section{Effective spin squeezing in the presence of a drive}
\label{sec:effective_squeezing_drive}

Suppose we have a Hamiltonian of the form
\begin{align}
  H = H_0 + V,
  &&
  H_0 = - \xi\v S\c\v S,
  &&
  V = - \bar h S_z - \Omega S_x - \sum_n b_n s_n^z,
\end{align}
where the effective fields $b_n$ are mean-zero, and we consider
initial states in the ground-state manifold $\G_0$ of $H_0$ (i.e.~with
total spin $S=N/2$).  If the largest eigenvalue of $V$ is smaller in
magnitude than half of the collective spin gap $N\xi$, or the energy
gap under $H_0$ between $\G_0$ and its orthogonal complement $\E_0$,
then we can formally develop a perturbative treatment of the action of
$V$ on $\G_0$.  Such a treatment yields an effective Hamiltonian
$H_{\t{eff}}=\sum_pH_{\t{eff}}^{(p)}$ on $\G_0$, where
$H_{\t{eff}}^{(p)}$ is order $p$ in $V$.  Letting $\P_0$ ($\Q_0$) be a
projector onto $\G_0$ ($\E_0$) and $X$ denote any operator on
$\G_0\cup\E_0$ (i.e.~the entire Hilbert space), we define the
superoperators
\begin{align}
  \D X \equiv \P_0 X \P_0 + \Q_0 X \Q_0,
  &&
  \O X \equiv \P_0 X \Q_0 + \Q_0 X \P_0,
\end{align}
which select the diagonal ($\D$) and off-diagonal ($\O$) parts of $X$
with respect to $\G_0$ and $\E_0$, and
\begin{align}
  \L X \equiv \sum_{\alpha,\beta}
  \f{\op{\alpha}\O X\op{\beta}}{E_\alpha-E_\beta},
  &&
  \t{where}
  &&
  H_0 = \sum_\alpha E_\alpha \op\alpha.
\end{align}
The first few terms in the expansion of the effective Hamiltonian are
then, as derived in ref.~\cite{bravyi2011schrieffer},
\begin{align}
  H_{\t{eff}}^{(0)} = \P_0 H_0 \P_0,
  &&
  H_{\t{eff}}^{(1)} = \P_0 V \P_0,
  &&
  H_{\t{eff}}^{(2)} = -\f12 \P_0 \sp{\O V,\L V} \P_0.
\end{align}
Letting $\B_0\p{\E_0}$ denote an eigenbasis of $H_0$ for $\E_0$ and
adding a constant shift to $H_0$ such that $E_\psi=0$ for
$\ket\psi\in\G_0$, we additionally define
\begin{align}
  \I \equiv \sum_{\ket\alpha\in\B_0\p{\E_0}} \f{\op\alpha}{E_\alpha},
\end{align}
which sums over projections onto excited states with corresponding
energetic suppression factors, and expand
\begin{align}
  \L X = \O\p{\L X}
  = \Q_0 \L X \P_0 + \P_0 \L X \Q_0
  = \I X \P_0 - \P_0 X \I.
\end{align}
The expression for $H_{\t{eff}}^{(2)}$ then simplifies to
\begin{align}
  H_{\t{eff}}
  = -\f12 \P_0 \p{\sp{\O V, \I V \P_0} - \sp{\O V, \P_0 V \I}} \P_0
  = -\P_0 V \I V \P_0.
\end{align}
\note{[MP: much of the text above is essentially copied from an
  appendix of the theory paper for multi-body interactions.  As soon
  as that paper gets posted on arXiv, we can cite it instead.]}  It
follows that
\begin{align}
  H_{\t{eff}}^{(0)} = 0,
  &&
  H_{\t{eff}}^{(1)} = -\bar h S_z - \Omega S_x,
\end{align}
and, in terms of the $S=N/2$ Dicke states $\ket{m}\equiv\ket{N/2,m}$
and the $S=N/2-1$ spin-wave states $\ket{mk}\equiv\ket{N/2-1,mk}$,
\begin{align}
  H_{\t{eff}}^{(2)}
  = - \f1{N\xi} \sum_{m,k} \abs{\sum_nb_n\bk{mk|s_n^z|m}}^2 \op{m},
  \label{eq:general_H_eff_2}
\end{align}
where we used the fact that the collective spin operators $S_j$
conserve total spin $S$, while the individual spin operators $s_n^z$
conserve the projection $m$ of collective spin onto a quantization
axis.  The second-order effective Hamiltonian $H_{\t{eff}}^{(2)}$ in
\eqref{eq:general_H_eff_2} is identical to that derived in Section
\ref{sec:OAT}, which implies
\begin{align}
  H_{\t{eff}}^{(2)} = \chi S_z^2,
  &&
  \chi \equiv \f{\tilde h^2}{N\p{N-1}\xi},
  &&
  \tilde h^2 \equiv \f1N \sum_n b_n^2,
\end{align}
and in turn
\begin{align}
  H_{\t{eff}} = -\bar h S_z - \Omega S_x + \chi S_z^2.
\end{align}

\bibliography{\jobname}

\end{document}

### 2018-07-09 multi_body_interactions/papers/theory/paper.tex
add multi-body paper

\documentclass[preprint,showkeys,nofootinbib]{revtex4-1}

% linking references
\usepackage{hyperref}
\hypersetup{
  breaklinks=true,
  colorlinks=true,
  linkcolor=blue,
  urlcolor=cyan,
}

% general physics / math packages and commands
\usepackage{physics,amsmath,amssymb,braket,dsfont,bm}
\renewcommand{\t}{\text} % text in math mode
\newcommand{\f}{\dfrac} % shorthand for fractions
\newcommand{\p}[1]{\left(#1\right)} % parenthesis
\renewcommand{\sp}[1]{\left[#1\right]} % square parenthesis
\renewcommand{\set}[1]{\left\{#1\right\}} % curly parenthesis
\newcommand{\bk}{\braket} % shorthand for braket
\newcommand{\g}{\text{g}}
\newcommand{\e}{\text{e}}
\renewcommand{\c}{\hat{c}}
\newcommand{\n}{\hat{n}}
\newcommand{\A}{\mathcal{A}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\G}{\mathcal{G}}
\renewcommand{\H}{\mathcal{H}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\K}{\mathcal{K}}
\renewcommand{\L}{\mathcal{L}}
\newcommand{\M}{\mathcal{M}}
\newcommand{\N}{\mathcal{N}}
\renewcommand{\O}{\mathcal{O}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\Q}{\mathcal{Q}}
\renewcommand{\S}{\mathcal{S}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\1}{\mathds{1}}

% "left vector" arrow; requires tikz package
\newcommand{\lvec}[1]
{\reflectbox{\ensuremath{\vec{\reflectbox{\ensuremath{#1}}}}}}

% figures
\usepackage{graphicx} % for figures
\usepackage{grffile} % help latex properly identify figure extensions
\graphicspath{{./figures/}} % set path for all figures
\usepackage[caption=false]{subfig} % subfigures (via \subfloat[]{})

% inline lists
\usepackage[inline]{enumitem}
\setlist[enumerate,1]{label={(\roman*)}}

% for feynman diagrams
\usepackage{tikz,tikz-feynman}
\tikzset{
  baseline = (current bounding box.center)
}
\tikzfeynmanset{
  compat = 1.1.0,
  every feynman = {/tikzfeynman/small}
}
\newcommand{\shrink}[1]{\scalebox{0.8}{#1}} % for smaller diagrams

% color definitions (used in a figure)
\usepackage{xcolor}
\definecolor{lightblue}{RGB}{31,119,180}
\definecolor{orange}{RGB}{255,127,14}
\definecolor{green}{RGB}{44,160,44}
\definecolor{lightred}{RGB}{214,39,40}

% proper coloring inside math environment
\makeatletter
\def\mathcolor#1#{\@mathcolor{#1}}
\def\@mathcolor#1#2#3{
  \protect\leavevmode
  \begingroup
    \color#1{#2}#3
  \endgroup
}
\makeatother
\newcommand{\bmu}{\mathcolor{lightblue}{\mu}}
\newcommand{\onu}{\mathcolor{orange}{\nu}}
\newcommand{\grho}{\mathcolor{green}{\rho}}
\newcommand{\re}{\mathcolor{lightred}{\text{e}}}

% leave a note in the text, visible in the compiled document
\newcommand{\note}[1]{\textcolor{red}{#1}}

% for strikeout text
% normalem included to prevent underlining titles in the bibliography
\usepackage[normalem]{ulem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\title{Effective multi-body SU($N$)-symmetric interactions of
  ultracold fermionic atoms on a 3-D lattice}

\author{M A Perlin}
\email{mika.perlin@gmail.com}
\author{A M Rey}
\affiliation{JILA, National Institute of Standards and Technology and
  University of Colorado, 440 UCB, Boulder, Colorado 80309, USA}
\affiliation{Center for Theory of Quantum Matter, 440 UCB, Boulder,
  Colorado 80309, USA}
\affiliation{Department of Physics, University of Colorado, 390 UCB,
  Boulder, Colorado 80309, USA}

\begin{abstract}
  Rapid advancements in the experimental capabilities with ultracold
  alkaline-earth-like atoms (AEAs) bring to a surprisingly near term
  the prospect of performing quantum simulations of spin models and
  lattice field theories exhibiting SU($N$) symmetry.  Motivated in
  particular by recent experiments preparing high density samples of
  strongly interacting ${}^{87}$Sr atoms in a three-dimensional
  optical lattice, we develop a low-energy effective theory of
  fermionic AEAs which exhibits emergent SU($N$)-symmetric multi-body
  interactions, where $N$ is the number of atomic nuclear spin levels.
  We fully characterize the low-lying excitations in our effective
  theory, and compare predictions of many-body interaction energies
  with direct measurements of many-body excitation spectra in an
  optical lattice clock.  Our work makes the first step in enabling a
  controlled, bottom-up experimental investigation of multi-body
  SU($N$) physics.
\end{abstract}

\keywords{\it ultracold atoms, optical lattices, SU(N) magnetism}

\maketitle

\tableofcontents

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:introduction}

% Rapid advancements in the experimental capabilities with ultracold
% alkaline-earth-like atoms (AEAs) \cite{fukuhara2007degenerate,
% stellmer2009boseeinstein, deescobar2009boseeinstein,
% tey2010doubledegenerate, desalvo2010degenerate, taie2012su,
% martin2013quantum, cappellini2014direct, scazza2014observation,
% zhang2014spectroscopic, cazalilla2014ultracold} bring to a
% surprisingly near term the prospect of performing quantum
% simulations of spin models and lattice field theories exhibiting
% SU($N$) symmetry \cite{cazalilla2009ultracold, hermele2009mott,
% hermele2011topological, gorshkov2010twoorbital, banerjee2013atomic,
% cazalilla2014ultracold}.  AEAs have ground and metastable
% (electronic) orbital states\note{[I want to briefly discuss how we
% present this]} ${}^1S_0$ and ${}^3P_0$, often referred to as the
% ``clock'' states due to their use in optical lattice clocks
% \cite{derevianko2011colloquium, katori2011optical}, with
% ultra-narrow (millihertz) optical transition linewidths and
% associated $\sim10^2$ second excited-state lifetimes.  Due to the
% lack of electronic orbital angular momentum in the ${}^1S_0$ and
% ${}^3P_0$ states, AEAs exhibit decoupled orbital and nuclear spin
% degrees of freedom.  For atoms with total nuclear spin $I$ and
% $N=2I+1$ nuclear spin levels, this decoupling leads to
% nuclear-spin-conserving SU($N$)-symmetric interactions governed
% entirely by orbital-state parameters \cite{cazalilla2009ultracold,
% taie2012su, zhang2014spectroscopic}.

Fermionic alkaline-earth atoms (AEAs), in addition to other atoms such
as ytterbium (Yb) sharing similar electronic structure, are currently
the building blocks of the most precise atomic clocks in the world
\cite{derevianko2011colloquium, katori2011optical, ludlow2015optical}.
The atoms have a unique, ultra-narrow optical transition between
metastable ${}^1S_0$ and ${}^3P_0$ electronic orbital states, i.e.~the
``clock states'', that allows for coherence times which can exceed 100
seconds \cite{porsev2004hyperfine, santra2004properties}.
Furthermore, AEAs can be trapped in fully controllable optical lattice
potentials and interrogated with ultra-stable lasers that can resolve
and probe their rich hyperfine spectra, consisting of $N$ different
nuclear spin levels with $N$ as large as 10 in strontium (${}^{87}$Sr)
and 6 in ytterbium (${}^{173}$Yb).

In 2015 the ${}^{87}$Sr optical lattice clock (OLC) at JILA, operated
in a one-dimensional (1-D) lattice at microkelvin temperatures,
achieved a total fractional uncertainty of $2\times10^{-18}$
\cite{bloom2014optical, nicholson2015systematic}.  Last year (2017), a
new generation of OLCs became operational at JILA, interrogating a
Fermi degenerate gas of ${}^{87}$Sr atoms in a 3-D lattice at
nanokelvin temperatures \cite{campbell2017fermidegenerate}.  All of
these atoms' degrees of freedom, including the electronic orbital,
nuclear spin, and motional states, can be fully controlled in a 3-D
lattice with high fidelity \cite{daley2008quantum,
  gorshkov2009alkalineearthmetal, daley2011quantum,
  daley2011statedependent}.  With frequency measurements reaching the
$10^{-19}$ fractional uncertainty level, the new OLCs are thus
enabling an exciting opportunity to probe, for the first time, quantum
dynamics with sub-millihertz spectral resolution
\cite{campbell2017fermidegenerate}.

A wonderful consequence of the efforts to build better clocks is the
development of highly controllable quantum simulators of many-body
systems in the strongly-interacting regime, where inter-particle
interactions set the largest energy scale relevant for system dynamics
\cite{swallows2011suppression, lemke2011wave,
  campbell2017fermidegenerate}.  The marriage between precision clock
spectroscopy and quantum many-body physics \cite{taie2012su,
  martin2013quantum, scazza2014observation, cappellini2014direct,
  zhang2014spectroscopic, rey2014probing} has an enormous potential to
enable novel explorations of physics for the same reason that makes
AEAs such remarkable time-keepers.  Specifically, due to the lack of
electronic orbital angular momentum in the ${}^1S_0$ and ${}^3P_0$
states, AEAs exhibit decoupled orbital and nuclear spin degrees of
freedom.  For atoms with $N$ nuclear spin levels, this decoupling
leads to nuclear-spin-conserving SU($N$)-symmetric interactions
governed entirely by orbital-state parameters
\cite{cazalilla2009ultracold, taie2012su, zhang2014spectroscopic}.

The presence of this exotic SU($N$) symmetry in a highly controllable
experimental platform opens the door to experimental studies of
e.g.~the SU($N$) Heisenberg model, whose phase diagram is believed to
exhibit features such as a chiral spin liquid (CSL) phase with
topological order and fractional statistics \cite{hermele2009mott,
  hermele2011topological, chen2016syntheticgaugefield}.  In addition
to illuminating open questions in our understanding of the fractional
quantum Hall effect and unconventional superconductivity
\cite{lee2006doping, lee2008high, gong2014emergent}, the CSL can
support non-Abelian excitations which allow for universal topological
quantum computation \cite{freedman2004class, hermele2011topological}.
Harnessing the SU($N$)-symmetric interactions of AEAs might also
enable the simulation of various lattice gauge theories
\cite{wieseu.-j.2013ultracold, zohar2016quantum}, some of which share
important qualitative features with quantum chromodynamics such as
few-body bound states and confinement \cite{banerjee2013atomic,
  rico2018nuclear}.  These direct, quantum simulations have an
extraordinary potential to provide novel insights by circumventing
severe sign problems which plague classical simulations of strongly
interacting fermionic systems \cite{wieseu.-j.2013ultracold,
  wu2003exact}.

% One of the major drivers of experimental capabilities with AEAs is
% the development of optical lattice clocks
% \cite{derevianko2011colloquium, katori2011optical}.  These clocks
% traditionally operate in a one-dimensional (1-D) geometry with
% harmonic confinement of atoms in the plane orthogonal to the lattice
% axis \cite{derevianko2011colloquium, hinkley2013atomic,
% bloom2014optical}.  By operating at higher atomic densities, one can
% simultaneously interrogate a larger number of atoms and thereby
% increase the signal-to-noise ratio of the clock.  With the extreme
% sensitivities accessible to state-of-the-art optical lattice clocks,
% however, 1-D clock experiments with ${}^{87}$Sr and ${}^{171}$Yb
% detect density-dependent fractional frequency shifts at the
% $10^{-16}$ level \cite{campbell2009probing, lemke2009spin}.  For the
% purpose of timekeeping, these frequency shifts amount to an
% undesirable source of systematic error, but they introduced a new,
% exciting paradigm for the field of quantum simulation: a marriage
% between precision clock spectroscopy and quantum many-body physics
% \cite{taie2012su, scazza2014observation, martin2013quantum,
% cappellini2014direct, zhang2014spectroscopic, rey2014probing}.
% Somewhat counter-intuitively, one of the most promising strategies
% for improving the interaction-limited 1-D clock design has been to
% use higher dimensional lattices to generate tighter on-site
% confinement of atoms, thereby {\it increasing} the strength of
% inter-atomic interactions and making evolution into interacting
% many-body states energetically unfavorable
% \cite{swallows2011suppression, lemke2011wave,
% campbell2017fermidegenerate}.  A wonderful side-effect of this
% strategy for building better clocks is the development of a highly
% controllable quantum many-body system in the strongly-interacting
% regime, where inter-atomic interactions set the largest energy scale
% relevant for system dynamics.

In this work, we investigate the first experimental capabilities with
ultracold fermionic AEAs to prepare high-density samples in a 3-D
lattice with multiple occupation of individual lattice sites
\cite{goban2018emergence}.  Specifically, we consider the preparation
of isolated few-body systems in the deep-lattice limit, and carry out
a bottom-up investigation of emergent multi-body interactions on
multiply-occupied lattice sites.  These multi-body interactions appear
in a low-energy effective theory of the atoms, and inherit the SU($N$)
symmetry of their bare, pair-wise interactions, thereby enabling
experimental studies of multi-body SU($N$) physics through the
exquisite capabilities with OLCs.

Though effective multi-body interactions have previously been studied
in the context of harmonically \cite{johnson2012effective,
  yin2014universal} and lattice-confined \cite{johnson2009effective}
neutral bosons prepared in a single hyperfine state, our work deals
for the first time with fermionic atoms with internal degrees of
freedom and multiple collisional parameters.  Nonetheless, due to the
SU($N$) symmetry of these collisions, we are able to find a simple way
to express effective multi-body Hamiltonians and fully characterize
their low-lying many-body energy eigenstates.  While some past work
has detected experimental signatures of multi-body interactions in the
form of quantum phase revivals \cite{will2010timeresolved}, we perform
a direct comparison of the many-body interaction energies predicted by
our low-energy effective theory with experimental measurements of
density-dependent atomic energy spectra performed in
ref.~\cite{goban2018emergence}, similarly to the measurements of
bosons performed in ref.~\cite{mark2011precision}.

The remainder of this paper is structured as follows.  In section
\ref{sec:overview} we provide an overview of the relevant one- and
two-body physics of ultracold atoms on a lattice, and summarize our
main result of effective multi-body interactions between these atoms.
In section \ref{sec:low_energy}, we discuss our method for deriving a
low-energy effective theory, provide a perturbative expansion for the
net effective Hamiltonian, and compute all $M$-body Hamiltonians
through third order in the low-energy effective theory.  We then
analyze the low-lying excitations of the effective theory in section
\ref{sec:spectra}, comparing spectral predictions with experimental
measurements, and study the orbital-state dynamics of nuclear spin
mixtures interrogated via Rabi spectroscopy.  Finally, we summarize
and conclude our findings in section \ref{sec:summary}, and provide
some discussion of future outlooks.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Overview}
\label{sec:overview}

Consider a collection of fermionic atoms with total nuclear spin $I$
and two atomic orbital states loaded into a translationally invariant
3-D optical lattice.  While an external trapping potential will
generally break discrete translational symmetry of the lattice, any
background inhomogeneity can be made negligible by spectroscopically
addressing a sufficiently small region of the lattice
\cite{goban2018emergence}.  Throughout this paper, we work strictly in
the deep-lattice regime with negligible tunneling between lattice
sites.  We also assume a so-called ``magic-wavelength'' lattice for
which both atomic orbital states experience identical lattice
potentials \cite{ye2008quantum}.  The single-particle Hamiltonian of
the atoms can then be written in the form
\begin{align}
  H_0 = \sum_{i,n,\mu,s} E_n \c_{in\mu s}^\dag \c_{in\mu s},
  \label{eq:H_0}
\end{align}
where $\c_{in\mu s}$ is a fermionic operator which annihilates a
single atom on lattice site $i\in\mathbb{Z}^3$ in motional state
$n\in\mathbb{N}_0^3$ with nuclear spin $\mu\in\set{-I,I+1,\cdots,I}$
(i.e.~projected onto a quantization axis) and orbital state
$s\in\set{\g,\e}$; and $E_n$ is the energy of a single atom on site
$i$ in motional state $n$.  In a harmonic trap approximation we would
have $E_n=\p{3/2+n_x+n_y+n_z}\omega$ for an on-site angular trap
frequency $\omega$ (with the reduced Planck constant $\hbar=1$
throughout this paper), but in general the aharmonicity of the lattice
potential will cause a non-negligible shift in motional state
energies.

In the absence of hyperfine coupling, as when addressing the spinless
${}^1S_0$ ($\g$) and ${}^3P_0$ ($\e$) orbital states, interactions
between any two atoms are governed by their orbital states alone, and
are therefore characterized by four scattering lengths $a_X$ with
$X\in\set{\g\g,\e\g^-,\e\g^+,\e\e}$, where the $+$ ($-$) superscript
denotes symmetrization (anti-symmetrization) of a two-body orbital
state under particle exchange.  In the low-energy limit, we can write
the bare two-body interaction Hamiltonian in the form
\cite{gorshkov2010twoorbital}
\begin{align}
  H_{\t{int}} = \sum_{\substack{\mu<\nu\\s}}
  G_s \int d^3x~ \hat\rho_{\mu s} \hat\rho_{\nu s}
  + G_+ \sum_{\mu,\nu}
  \int d^3x~ \hat\rho_{\mu,\e} \hat\rho_{\nu,\g}
  + G_- \sum_{\mu,\nu} \int d^3x~
  \hat\psi_{\mu,\e}^\dag \hat\psi_{\nu,\g}^\dag
  \hat\psi_{\nu,\e} \hat\psi_{\mu,\g},
\end{align}
where $\hat\psi_{\mu s}$ is a fermionic field operator for atoms with
nuclear spin $\mu$ and orbital state $s$;
$\hat\rho_{\mu s}\equiv\hat\psi_{\mu s}^\dag\hat\psi_{\mu s}$ is an
atomic density field operator; and the coupling constants $G_X$ are
defined in terms of the scattering lengths by
\begin{align}
  G_{s=\g,\e} \equiv \f{4\pi}{m_A}~ a_{ss}, &&
  G_\pm \equiv \f{2\pi}{m_A}~ \p{a_{\e\g+} \pm a_{\e\g-}},
  \label{eq:couplings}
\end{align}
where $m_A$ is the mass of a single atom.  Defining the coupling
tensor
\begin{align}
  G^{qr}_{st} \equiv \left\{
    \begin{array}{ll}
      G_q & ~ q = r = s = t \\
      G_+ & ~ q \ne r ~ \t{and} ~ (q,r) = (s,t) \\
      G_- & ~ q \ne r ~ \t{and} ~ (q,r) = (t,s) \\
      0 & ~ \t{otherwise}
    \end{array}\right.,
  \label{eq:coupling_tensor}
\end{align}
where $q,r,s,t\in\set{\g,\e}$ are orbital state indices, we can
alternately write the bare two-body interaction Hamiltonian in the
more compact form
\begin{align}
  H_{\t{int}}
  = \f12 \sum_{\substack{q,r,s,t\\\mu,\nu}} G^{qr}_{st} \int d^3x~
  \hat\psi_{\mu s}^\dag \hat\psi_{\nu t}^\dag
  \hat\psi_{\nu r} \hat\psi_{\mu q}.
  \label{eq:H_int_fields}
\end{align}
For nuclear spins $\mu,\nu$, the coupling tensor $G^{qr}_{st}$ gives
the coupling constant between the two-atom states
$(\mu,q)+(\nu,r)\leftrightarrow(\mu,s)+(\nu,t)$.

We now expand the field operators $\hat\psi_{\mu s}$ in the Wannier
basis for a 3-D lattice, such that
$\hat\psi_{\mu s}(x) = \sum_{i,n} \phi_{in}(x) \c_{in\mu s}$ with
spatial wavefunctions $\phi_{in}$ and fermionic annihilation operators
$\c_{in\mu s}$ indexed by lattice sites $i$ and motional states $n$.
Invoking the tight-binding approximation, we assume that the spatial
overlap integral in \eqref{eq:H_int_fields} is negligible unless all
wavefunctions are localized at the same lattice site; we discuss the
breakdown of this approximation and its consequences for our
low-energy effective theory in Appendix \ref{sec:error}.  The relevant
spatial overlap integral is then
\begin{align}
  K^{k\ell}_{mn}
  \equiv \int d^3x~ \phi_{im}^* \phi_{in}^* \phi_{i\ell} \phi_{ik},
  \label{eq:K_klmn}
\end{align}
which for a lattice with discrete translational invariance is
independent of the lattice site $i$.  The two-body interaction
Hamiltonian can be written in terms of this overlap integral as
\begin{align}
  H_{\t{int}}
  = \f12 \sum_{\substack{i,k,\ell,m,n\\q,r,s,t\\\mu,\nu}}
  K^{k\ell}_{mn} G^{qr}_{st}
  \c_{im\mu s}^\dag \c_{in\nu t}^\dag \c_{i\ell\nu r} \c_{ik\mu q}
  \equiv \f12 \sum K^{k\ell}_{mn} G^{qr}_{st}
  \c_{m\mu s}^\dag \c_{n\nu t}^\dag \c_{\ell\nu r} \c_{k\mu q},
  \label{eq:H_int}
\end{align}
where for brevity we will henceforth suppress the identical site index
($i$) on all operators, and implicitly sum over all indices in a
summand which do not have a fixed value.  We may also at times
suppress motional state indices on the overlap integral
$K^{k\ell}_{mn}$, in which case the suppressed indices are implicitly
zero (corresponding to a motional ground state); i.e.
\begin{align}
  K^{\ell m}_n \equiv K^{\ell m}_{n,0},
  &&
  K^m_n \equiv K^{m,0}_{n,0},
  &&
  K_{mn} \equiv K^{0,0}_{mn},
  &&
  K_n \equiv K^{0,0}_{n,0},
  &&
  K \equiv K^{0,0}_{0,0}.
  \label{eq:K}
\end{align}
For simplicity, we will also generally work in a gauge for which all
two-body overlap integrals are real, such that
$K^{k\ell}_{mn}={K^{k\ell}_{mn}}^*=K^{mn}_{k\ell}$.  This choice is
always possible due to the analytic properties of Wannier orbitals
$\phi_{in}$ \cite{kohn1959analytic}.

We are interested in the ground and low-lying orbital eigenstates of
several interacting atoms on a single lattice site, as such states can
be readily prepared and coherently addressed in current experiments
with ultracold atoms \cite{goban2018emergence}.  Although these
experiments operate at temperatures sufficiently low for negligible
thermal occupation of excited motional states, simply neglecting
excited single-particle motional states fails to reproduce the
observed spectra of multiply-occupied lattice sites
\cite{goban2018emergence}.  The reason for this failure lies in the
fact that the bare two-body Hamiltonian $H_{\t{int}}$ in
\eqref{eq:H_int} is not diagonal in the basis of single-particle
motional states.  Even at zero temperature, therefore, the presence of
excited motional states modifies many-body atomic energy spectra when
off-diagonal elements of $H_{\t{int}}$ approach the scale of low-lying
single-particle motional state energy gaps.  To address this problem,
we develop a low-energy effective theory for ultracold two-level
fermionic atoms on a lattice.  As our work is motivated by progress in
experiments which initialize all atoms in identical (i.e.~ground)
orbital and motional states, to simplify our theory we assume that any
$N$ atoms on a single lattice site occupy $N$ nuclear spin states.
Multiple occupation of a single nuclear spin state on the same lattice
site is initially forbidden by fermionic statistics, and subsequently
cannot be violated in the absence of tunneling and hyperfine coupling.

Our low-energy effective theory exhibits SU($N$)-symmetric multi-body
interactions, such that the effective interaction Hamiltonian can be
written in the form
\begin{align}
  H_{\t{int}}^{\t{eff}} = \sum_{M=2}^{2I+1} H_M,
\end{align}
where $H_M$ is an $M$-body Hamiltonian, $I$ is the total nuclear spin
of each atom (e.g.~$I=9/2$ for ${}^{87}$Sr), and the sum ends at
$2I+1$ because this is the largest number of atoms which may initially
occupy a single lattice site.  Focusing on the low-lying orbital state
excitations of this theory, we find that we can express multi-body
Hamiltonians in the simple form
\begin{align}
  H_M = \sum_{\abs{\set{\mu_j}}=M}
  \p{U_{M,\g} \prod_{\alpha=1}^M \n_{\mu_\alpha,\g}
    + U_{M,+} \n_{\mu_1,\e} \prod_{\alpha=2}^M \n_{\mu_\alpha,\g}
    + U_{M,-} \c_{\mu_1,\g}^\dag \c_{\mu_2,\e}^\dag
    \c_{\mu_2,\g} \c_{\mu_1,\e} \prod_{\alpha=3}^M \n_{\mu_\alpha,\g}},
  \label{eq:H_M_preview}
\end{align}
where $\c_{\mu s}\equiv\c_{0,\mu s}$ is a fermionic annihilation
operator for an atom in the motional ground state with nuclear spin
$\mu$ and orbital state $s$;
$\n_{\mu s}\equiv \c_{\mu s}^\dag \c_{\mu s}$ is a number operator;
and the sum is performed over all choices of nuclear spins $\mu_j$
with $j=1,\cdots,M$ for which all $\mu_j$ are distinct, or
equivalently all choices of $\mu_j$ for which the set $\set{\mu_j}$
contains $M$ elements, for a total of ${2I+1\choose M}\times\p{M!}$
nuclear spin combinations.  The coefficients $U_X$ can be expressed in
terms of the coupling constants of the bare two-body interaction
Hamiltonian (see Appendix \ref{sec:U_X}).  The key feature of the
$M$-body interactions in \eqref{eq:H_M_preview} is that they take a
form identical to the two-body interactions between motional
ground-state atoms, but with the addition of $M-2$ spectator atoms.
This form is a direct consequence of the SU($N$) symmetry of
underlying two-body interactions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Low-energy effective theory}
\label{sec:low_energy}

As the bare two-body Hamiltonian $H_{\t{int}}$ is not diagonal with
respect to single-particle motional states, the problem of determining
many-body spectra and eigenstates nominally involves all motional
degrees of freedom of the atoms involved.  If we are interested in
addressing internal atomic states at zero temperature, however, in
principle we need only to consider the motional ground state
associated with each many-body internal atomic state.  We can then
ignore all excited many-body motional states, as they will be neither
thermally occupied nor externally interrogated.  This procedure makes
the many-body problem more tractable by greatly reducing the
dimensionality of our target Hilbert space.

We denote the space of all internal states of atoms in
single-particle, non-interacting (multi-particle, interacting)
motional ground states by $\H_{\t{ground}}^{\t{single}}$
($\H_{\t{ground}}^{\t{multi}}$).  For sufficiently weak interactions,
one can construct a unitary transformation $U$ acting on the full
many-body Hilbert space which rotates $\H_{\t{ground}}^{\t{multi}}$
into $\H_{\t{ground}}^{\t{single}}$ \cite{bravyi2011schrieffer}.  The
unitary $U$ can then be used to construct an {\it effective
  Hamiltonian} $H_{\t{int}}^{\t{eff}}$ on
$\H_{\t{ground}}^{\t{single}}$ that reproduces the spectrum of
$H_{\t{int}}$ on $\H_{\t{ground}}^{\t{multi}}$, namely
$H_{\t{int}}^{\t{eff}}=UH_{\t{int}}U^\dag$.  This method for
constructing an effective theory is commonly known as the
Schrieffer-Wolff transformation, named after the authors of its
celebrated application in relating the Anderson and Kondo models of
magnetic impurities in metals \cite{schrieffer1966relation}.

Using the machinery developed in ref.~\cite{bravyi2011schrieffer} for
performing a rotation between low-energy subspaces of a perturbed
(i.e.~interacting) and unperturbed (i.e.~non-interacting) Hamiltonian,
we derive an expansion for the effective interaction Hamiltonian
$H_{\t{int}}^{\t{eff}}$ in terms two-body interaction Hamiltonian
$H_{\t{int}}$ (see Appendix \ref{sec:eff_derivation}).  This expansion
takes the form
\begin{align}
  H_{\t{int}}^{\t{eff}} = \sum_{p\ge1} H_{\t{int}}^{(p)},
  \label{eq:H_int_eff}
\end{align}
where $H_{\t{int}}^{(p)}$ is order $p$ in $H_{\t{int}}$.  Letting
$\E_0$ denote the orthogonal complement of
$\H_{\t{ground}}^{\t{single}}$ (i.e.~$\E_0$ is the space of all states
with at least one atom in an excited motional state), $\B_0\p{\E_0}$
denote an eigenbasis of $\E_0$ with respect to the single-particle
Hamiltonian $H_0$, and $E_\alpha$ denote the motional energy (with
respect to $H_0$) of a state $\ket\alpha\in\E_0$ relative to the
corresponding motional ground-state energy, we define the operator
\begin{align}
  \I \equiv \sum_{\ket\alpha\in\B_0\p{\E_0}} \f{\op\alpha}{E_\alpha},
\end{align}
which sums over projections onto excited states with corresponding
energetic suppression factors.  The operator $\I$ together with the
projector $\P_0$ onto $\H_{\t{ground}}^{\t{single}}$ allows us
concisely write the first few terms in \eqref{eq:H_int_eff} as
\begin{align}
  H_{\t{int}}^{(1)} = \P_0 H_{\t{int}} \P_0,
  &&
  H_{\t{int}}^{(2)} = -\P_0 H_{\t{int}} \I H_{\t{int}} \P_0,
  \label{eq:H_int_1_2}
\end{align}
\begin{align}
  H_{\t{int}}^{(3)}
  = \P_0 H_{\t{int}} \I H_{\t{int}} \I H_{\t{int}} \P_0
  - \f12\sp{\P_0 H_{\t{int}} \P_0,
    \P_0 H_{\t{int}} \I^2 H_{\t{int}} \P_0}_+,
  \label{eq:H_int_3}
\end{align}
where $\sp{X,Y}_+\equiv XY+YX$.  The naive idea of applying
$H_{\t{int}}$ directly on the single-particle motional ground states
and neglecting all excited states is thus equivalent to truncating our
expansion for $H_{\t{int}}^{\t{eff}}$ at first order.  In addition to
this first order term, the expansion involves {\it effective
  corrections} to the action of $H_{\t{int}}$ on the single-particle
motional ground states (i.e.~on $\H_{\t{ground}}^{\t{single}}$) in the
form of higher-order terms with {\it intermediate} or {\it virtual
  occupation} of excited states, manifest in $\I$.

Substituting the definition of $\I$ into
\eqref{eq:H_int_1_2}-\eqref{eq:H_int_3} yields expressions that are
highly reminiscent of standard non-degenerate perturbation theory in
quantum mechanics, but which nonetheless exhibit crucial differences.
The first, and most obvious difference is that these expressions are
operator equations, and that the sums over virtual states are
performed over a basis for the orthogonal complement of
$\H_{\t{ground}}^{\t{single}}$, rather than a basis for the orthogonal
complement of a single state, as in non-degenerate perturbation
theory.  Second, the non-degeneracy condition in standard perturbation
theory is here elevated to a restriction on the magnitude of the
perturbation $H_{\t{int}}$ relative to the spectral gap $\Delta$ of
the non-interacting Hamiltonian $H_0$ between
$\H_{\t{ground}}^{\t{single}}$ and the excited subspace $\E_0$.
Specifically, the validity of \eqref{eq:H_int_1_2}-\eqref{eq:H_int_3}
is conditional only on $\norm{H_{\t{int}}}\le\Delta/2$, where
$\norm{X}\equiv\max_{\ket\psi\in\H}\sqrt{\bk{\psi|X^\dag X|\psi}}$ is
the operator norm, with no restrictions on spectral gaps or
degeneracies within $\H_{\t{ground}}^{\t{single}}$ \cite{davis1969new,
  bravyi2011schrieffer}.  Finally, the effective theory involves no
corrections to the many-body energy eigenstates.  Including such
corrections would invalidate the effective theory, which preserves the
eigenstates of the non-interacting Hamiltonian $H_0$ on
$\H_{\t{ground}}^{\t{single}}$.

As a last comment, we note that our chosen method for constructing an
effective Hamiltonian is distinct from adiabatic elimination methods
which are commonly used in the atomic physics and quantum optics
communities to develop effective theories for e.g.~the low-lying
levels of a Lambda system \cite{brion2007adiabatic,
  james2007effective, reiter2012effective, sanz2016adiabatic}.  Unlike
the perturbative, but exact Schieffer-Wolff transformation, adiabatic
elimination methods use approximations which rely on the fast dynamics
of excited states.  While generally reasonable, these approximations
must be made carefully to avoid potential problems with
self-consistency (see section 3 of ref.~\cite{brion2007adiabatic}),
and yield no obvious or straightforward means to compute effective
corrections beyond second order in the couplings between low- and
high-energy sectors of a Hilbert space \cite{james2007effective,
  reiter2012effective, sanz2016adiabatic}.  While at least one attempt
at systematically computing higher-order corrections in the framework
of adiabatic elimination has recently been made
\cite{sanz2016adiabatic}, the resulting expressions do not lend
themselves as nicely to analytical treatment, and were in any case
found by the authors to be equivalent to a Schrieffer-Wolff
transformation.

\subsection{Diagrammatic representation of effective Hamiltonians}

The form of the bare two-body interaction Hamiltonian $H_{\t{int}}$ in
\eqref{eq:H_int} motivates a diagrammatic representation of terms in
the effective Hamiltonian $H_{\t{eff}}$ in \eqref{eq:H_int_eff},
similarly to the diagrams used to represent elements of the scattering
matrix in field theory.  Effective $M$-body interaction terms at order
$p$ in $H_{\t{int}}$ can be represented by directed, weakly connected
graphs (i.e.~which are connected when ignoring edge direction)
containing $p$ internal vertices.  Each internal vertex and its
associated edges correspond respectively to a coupling constant and
the associated field operators in $H_{\t{int}}$.  An example 2-vertex
diagram representing an effective 3-body interaction term is provided
in figure \ref{fig:diagram}.  All diagrams are read from left to right
to construct a sequence of operators from right to left; the internal
vertices of a diagram are thus strictly ordered, with the $n$-th
internal vertex from the left corresponding to the $n$-th interaction
Hamiltonian $H_{\t{int}}$ from the right in \eqref{eq:H_int_1_2} or
\eqref{eq:H_int_3}.  Solid (dashed) lines represent field operators
acting on the lowest (arbitrary) motional states.  Spatial overlap
factors at each vertex are determined by the motional states of the
edges which connect to (i.e.~field operators associated with) that
vertex.  While it is possible to construct explicit rules for
determining the energetic suppression factors (i.e.~from $\I$) of the
term represented by a diagram, these factors are most easily
determined by examination of the effective Hamiltonians in
\eqref{eq:H_int_1_2} and \eqref{eq:H_int_3}.

\begin{figure}
  \centering
  \(
  \begin{tikzpicture}
    \begin{feynman}
      \vertex (v1);
      \vertex[above left = of v1] (f1) {$\bmu,\g$};
      \vertex[below left = of v1] (f2) {$\onu,\re$};
      \vertex[right = of v1] (vm);
      \vertex[right = of vm] (v2);
      \vertex[above = of vm] (f3) {$\bmu,\g$};
      \vertex[below = of vm] (f4) {$\grho,\g$};
      \vertex[below right = of v2] (f5) {$\grho,\re$};
      \vertex[above right = of v2] (f6) {$\onu,\g$};
      \vertex [left = 0.5em of v1] {$G^{\g\re}_{\g\re}$};
      \vertex [right = 0.5em of v2] {$G^{\re\g}_{\g\re}$};
      \diagram* {
        (f1) --[fermion, color = lightblue] (v1)
        --[fermion, color = lightblue] (f3),
        (f2) --[fermion, color = orange] (v1)
        --[charged scalar,
        edge label = {$\mathcolor{black}{n}
          \mathcolor{orange}{\nu}\mathcolor{black}{,\re}$},
        color = orange] (v2),
        (f4) --[fermion, color = green] (v2)
        --[fermion, color = green] (f5),
        (v2) --[fermion, color = orange] (f6), };
    \end{feynman}
  \end{tikzpicture}
  \begin{array}{ll}
    &= \P_0 \p{\f12 K_n G^{\re\g}_{\g\re}
      \c_{\onu,\g}^\dag \c_{\grho,\re}^\dag
      \c_{\grho,\g} \c_{n\onu,\re}}
      \f1{E_n}
      \p{\f12 K_n G^{\g\re}_{\g\re}
      \c_{\bmu,\g}^\dag \c_{n\onu,\re}^\dag
      \c_{\onu,\re} \c_{\bmu,\g}} \P_0 \\[1em]
    &= \f14 \f{K_n^2}{E_n}~ G^{\g\re}_{\g\re} G^{\re\g}_{\g\re}~
      \P_0 \c_{\bmu,\g}^\dag \c_{\onu,\g}^\dag \c_{\grho,\re}^\dag
      \c_{\grho,\g} \c_{\onu,\re} \c_{\bmu,\g} \P_0
  \end{array}
  \)
  \caption{\footnotesize An example second-order diagram and the
    corresponding three-body interaction term in $H_{\t{int}}^{(2)}$,
    with $n>0$ and $\c_{\mu s}\equiv\c_{0,\mu s}$.  For the sake of
    presentation, this diagram has colors associated with nuclear spin
    and orbital states, an arrow on each line to emphasize that they
    are directed left-to-right, and an explicit coupling constant
    written next to each vertex; we will generally not include these
    features, as they are not necessary to uniquely identify a
    diagram.  We will also drop explicit appearances of the
    ground-state projector $\P_0$ in our expressions, with the
    understanding that the low-energy effective theory implicitly
    addresses only single-atom motional ground states.}
  \label{fig:diagram}
\end{figure}

The diagram in figure \ref{fig:diagram} explicitly labels all edges
with indices of the corresponding field operators, but in general we
may suppress these indices, in which case the diagram includes a sum
over the suppressed indices.  These sums are performed over all
allowed values of the suppressed indices, with the restriction that
virtual states (i.e.~vertical slices of the diagram between internal
vertices) represented with dashed lines must have at least one
motional excitation.  While we include factors of $1/2$ from
$H_{\t{int}}$ as expressed in \eqref{eq:H_int} in the definition of a
diagram, in all but the two-body case these factors of $1/2$ will be
cancelled out by corresponding symmetry factors, i.e.~the appearance
of duplicate diagrams which are equal up to a relabeling of indices
(see Appendix \ref{sec:diagrams}).  The explicit signs and factor of
$1/2$ which appear in the effective Hamiltonians in
\eqref{eq:H_int_1_2} and \eqref{eq:H_int_3} are not included in the
definition of a diagram, and must be kept track of manually.

\subsection{Effective two-body interactions and renormalization}
\label{sec:two_body}

The effective two-body Hamiltonian in \eqref{eq:H_int_eff} has
contributions at all orders in the coupling constants, and can be
expanded in the form
\begin{align}
  H_2 = \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v);
        \vertex[above left = of v] (f1);
        \vertex[below left = of v] (f2);
        \vertex[above right = of v] (f3);
        \vertex[below right = of v] (f4);
        \diagram* {
          (f1) -- (v),
          (f2) -- (v),
          (v) -- (f3),
          (v) -- (f4) };
      \end{feynman}
    \end{tikzpicture}}
  - \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[right = of v1] (v2);
        \vertex[above right = of v2] (f3);
        \vertex[below right = of v2] (f4);
        \diagram* {
          (f1) -- (v1),
          (f2) -- (v1),
          (v2) -- (f3),
          (v2) -- (f4),
          (v1) --[scalar, half left] (v2),
          (v1) --[scalar, half right] (v2) };
      \end{feynman}
    \end{tikzpicture}}
  + \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[right = of v1] (v2);
        \vertex[right = of v2] (v3);
        \vertex[above right = of v3] (f3);
        \vertex[below right = of v3] (f4);
        \diagram* {
          (f1) -- (v1),
          (f2) -- (v1),
          (v3) -- (f3),
          (v3) -- (f4),
          (v1)
          --[scalar, half left] (v2)
          --[scalar, half left] (v3),
          (v1)
          --[scalar, half right] (v2)
          --[scalar, half right] (v3) };
      \end{feynman}
    \end{tikzpicture}}
  + \cdots \equiv \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex[blob] (v) {};
        \vertex[above left = of v] (f1);
        \vertex[below left = of v] (f2);
        \vertex[above right = of v] (f3);
        \vertex[below right = of v] (f4);
        \diagram* {
          (f1) -- (v),
          (f2) -- (v),
          (v) -- (f3),
          (v) -- (f4) };
      \end{feynman}
    \end{tikzpicture}},
  \label{eq:H_2_expansion}
\end{align}
where the blob on the right schematically represents the net effective
two-body interaction.  In general, the sums over excited states in the
loop diagrams of \eqref{eq:H_2_expansion} will diverge
\cite{johnson2012effective}, which implies the need to renormalize the
coupling constants in our theory.  Even if these diagrams do not
diverge, we would like to express the effective two-body interactions
in terms of net, physically observable two-body coupling constants as
on the right-hand side of \eqref{eq:H_2_expansion}, rather than in
terms of long sums at all orders of the bare coupling constants.  We
therefore renormalize our coupling constants by introducing
counter-terms $\tilde G_X$ into the effective interaction Hamiltonians
via $G_X\to G_X+\tilde G_X$.  The values of these counter-terms are
fixed by enforcing that $G_X$ is equal to the net two-body interaction
strength.  Diagrammatically, this renormalization condition is
\begin{align}
  \substack{
    \shrink{
      \begin{tikzpicture}
        \begin{feynman}
          \vertex (v);
          \vertex[above left = of v] (f1);
          \vertex[below left = of v] (f2);
          \vertex[above right = of v] (f3);
          \vertex[below right = of v] (f4);
          \diagram* {
            (f1) -- (v),
            (f2) -- (v),
            (v) -- (f3),
            (v) -- (f4) };
        \end{feynman}
      \end{tikzpicture}}
    \\ \O(G)}
  + \substack{
    \shrink{
      \begin{tikzpicture}
        \begin{feynman}
          \vertex[crossed dot] (v) {};
          \vertex[above left = of v] (f1);
          \vertex[below left = of v] (f2);
          \vertex[above right = of v] (f3);
          \vertex[below right = of v] (f4);
          \diagram* {
            (f1) -- (v),
            (f2) -- (v),
            (v) -- (f3),
            (v) -- (f4) };
        \end{feynman}
      \end{tikzpicture}}
    \\ \O(\tilde G)}
  - \substack{
    \shrink{
      \begin{tikzpicture}
        \begin{feynman}
          \vertex (v1);
          \vertex[above left = of v1] (f1);
          \vertex[below left = of v1] (f2);
          \vertex[right = of v1] (v2);
          \vertex[above right = of v2] (f3);
          \vertex[below right = of v2] (f4);
          \diagram* {
            (f1) -- (v1),
            (f2) -- (v1),
            (v2) -- (f3),
            (v2) -- (f4),
            (v1) --[scalar, half left] (v2),
            (v1) --[scalar, half right] (v2) };
        \end{feynman}
      \end{tikzpicture}}
    \\ \O(G^2)}
  + \cdots = \substack{
    \shrink{
      \begin{tikzpicture}
        \begin{feynman}
          \vertex (v);
          \vertex[above left = of v] (f1);
          \vertex[below left = of v] (f2);
          \vertex[above right = of v] (f3);
          \vertex[below right = of v] (f4);
          \diagram* {
            (f1) -- (v),
            (f2) -- (v),
            (v) -- (f3),
            (v) -- (f4) };
        \end{feynman}
      \end{tikzpicture}}
    \\ \O(G)},
  \label{eq:counter_term}
\end{align}
where we represent counter-terms by a crossed dot (i.e.~$\otimes$) at
a vertex. To leading order in the coupling constants, it follows that
the counter-terms are second order in the coupling constants,
i.e.~$\tilde G\sim G^2$.  This finding implies that counter-terms only
come into play at next-to-leading order (NLO) in the calculation of
any effective $M$-body interaction Hamiltonian.  By construction,
these counter-terms exactly cancel all terms beyond leading order in
\eqref{eq:H_2_expansion}, which implies that the effective two-body
interaction Hamiltonian is simply
\begin{align}
  H_2 = \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v);
        \vertex[above left = of v] (f1);
        \vertex[below left = of v] (f2);
        \vertex[above right = of v] (f3);
        \vertex[below right = of v] (f4);
        \diagram* {
          (f1) -- (v),
          (f2) -- (v),
          (v) -- (f3),
          (v) -- (f4), };
      \end{feynman}
    \end{tikzpicture}}
  = \f12 \alpha_2^{(1)} \sum_{\abs{\set{\mu,\nu}}=2} G^{qr}_{st}
  \c_{\mu s}^\dag \c_{\nu t}^\dag \c_{\nu r} \c_{\mu q},
  \label{eq:H_2}
\end{align}
where for consistency with existing literature
\cite{johnson2012effective} we define $\alpha_2^{(1)} \equiv K$ as the
overlap integral between two atoms occupying single-particle motional
ground states.

Before moving on to consider effective three-body interactions, there
are two comments we must make concerning the result in \eqref{eq:H_2}.
First, this effective two-body interaction takes the same form as the
bare two-body interaction in \eqref{eq:H_int}, but without excited
motional states, and with renormalized coupling constants in place of
bare ones.  Our choice of renormalization scheme, while convenient for
the analytical development of a low-energy effective theory, no longer
allows us to use the coupling constants $G_X$ as defined by the
free-space scattering lengths $a_X$ in \eqref{eq:couplings} to compute
interaction energies.  Instead, we must first compute the effective
coupling constants $G_X^{\t{lattice}}\p{\U}$ in a lattice, which now
depend on the lattice depth $\U$, and in turn use the effective
coupling constants to compute interaction energies.  We discuss the
calculation of effective coupling constants in Appendix
\ref{sec:renormalization}.

Second, the overview we provided in section \ref{sec:overview}, and
consequently our result in \eqref{eq:H_2}, neglects the effect of
momentum-dependent two-body scattering.  When the effective range of
momentum-dependent interactions is comparable to the scattering
lengths $a_X$, as is the case for ultracold ${}^{87}$Sr, these
interactions are third order in the coupling constants $G_X$.  Just as
the $\O\p{G^2}$ counter-terms do not affect $M$-body interactions
until next-to-leading order (NLO), the $\O\p{G^3}$ momentum-dependent
interactions do not come into play until next-next-leading order
(NNLO).  Given that we develop our low-energy effective theory through
third order in the coupling constants, these interactions will not
appear in any of our three- and four-body calculations, but they do
need to be considered in the calculation of pair-wise interaction
energies.  The primary interest of our work, however, concerns
effective $M$-body interactions for $M\ge3$; we therefore defer the
calculation of momentum-dependent two-body interactions to Appendix
\ref{sec:momentum_dependence}.

\subsection{Effective three-body interactions at second order}

At second order in the coupling constants, the expansion for
$H_{\t{int}}^{(2)}$ in \eqref{eq:H_int_1_2} gives rise to effective
three-body interactions.  The virtual state of three-body terms in
$H_{\t{int}}^{(2)}$ cannot have two atoms in excited motional states,
as otherwise the second application of $H_{\t{int}}$ in
$H_{\t{int}}^{(2)}$ would have to address both of theses atoms to
bring them back down to the ground state, resulting in a two-body
process as in the second diagram of \eqref{eq:H_2_expansion}.  All
second-order three-body terms must therefore have only one excited
atom in the virtual state, and take the form
\begin{align}
  \begin{tikzpicture}
    \begin{feynman}
      \vertex (v1);
      \vertex[above left = of v1] (f1) {$\mu r$};
      \vertex[below left = of v1] (f2) {$\nu s$};
      \vertex[right = of v1] (vm);
      \vertex[right = of vm] (v2);
      \vertex[above = of vm] (f3) {$\mu r'$};
      \vertex[below = of vm] (f4) {$\rho t$};
      \vertex[below right = of v2] (f5) {$\rho t'$};
      \vertex[above right = of v2] (f6) {$\nu s''$};
      \diagram* {
        (f1) -- (v1) -- (f3),
        (f2) -- (v1) --  [scalar, edge label=$n\nu s'$] (v2),
        (f4) -- (v2) -- (f5),
        (v2) -- (f6), };
    \end{feynman}
  \end{tikzpicture}
  \propto K_n^2 G^{rs}_{r's'} G^{s't}_{s''t'}
  \c_{\mu r'}^\dag \c_{\nu s''}^\dag \c_{\rho t'}^\dag
  \c_{\rho t} \c_{\nu s} \c_{\mu r}
  \label{eq:H_3_2_diagram}.
\end{align}
Unlike for the two-body diagram in \eqref{eq:H_2}, the explicit
factors of $1/2$ which appear in the bare two-body Hamiltonian
$H_{\t{int}}$ in \eqref{eq:H_int} are now cancelled out by symmetry
factors which account for duplicate diagrams; this cancellation will
generally occur for all connected $M$-body diagrams with $M>2$ (see
Appendix \ref{sec:diagrams}).  The net effective three-body
interaction Hamiltonian at second order is then given by the sum over
all diagrams of the form in \eqref{eq:H_3_2_diagram}, i.e.
\begin{align}
  H_3^{(2)} = - \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[right = 4em of v1] (v2);
        \vertex[above right = of v1] (f3);
        \vertex[below left = of v2] (f4);
        \vertex[below right = of v2] (f5);
        \vertex[above right = of v2] (f6);
        \diagram* {
          (f1) -- (v1) -- (f3),
          (f2) -- (v1) --[scalar] (v2),
          (f4) -- (v2) -- (f5),
          (v2) -- (f6), };
      \end{feynman}
    \end{tikzpicture}}
  = -\alpha_3^{(2)} \sum_{\abs{\set{\mu,\nu,\rho}}=3}
  G^{rs}_{r's'} G^{s't}_{s''t'}
  \c_{\mu r'}^\dag \c_{\nu s''}^\dag \c_{\rho t'}^\dag
  \c_{\rho t} \c_{\nu s} \c_{\mu r},
  \label{eq:H_3_2}
\end{align}
where $\alpha_3^{(2)} \equiv \sum_{n>0} K_n^2/E_n$, and the preceding
negative sign is as prescribed by $H_{\t{int}}^{(2)}$ in
\eqref{eq:H_int_1_2}.

\subsection{Effective three-body interactions at third order}

The third-order effective interaction Hamiltonian $H_{\t{int}}^{(3)}$
in \eqref{eq:H_int_3} contains both three- and four-body terms.  To
compactly enumerate and evaluate all three-body diagrams at third
order, we introduce an expanded coupling tensor
\begin{align}
  G^{\mu q;\nu r}_{\rho s;\sigma t} \equiv \left\{
    \begin{array}{ll}
      G^{qr}_{st} & ~ \p{\mu,\nu} = \p{\rho,\sigma}
      \\
      - G^{qr}_{ts} & ~ \p{\mu,\nu} = \p{\sigma,\rho}
      \\
      0 & ~ \t{otherwise}
    \end{array}\right.
  \label{eq:G_expanded}
\end{align}
for more general
$\p{\mu,q}+\p{\nu,r}\leftrightarrow\p{\rho,s}+\p{\sigma,t}$ coupling
induced by terms proportional to
$\c_{\mu q}^\dag \c_{\nu r}^\dag \c_{\sigma t} \c_{\rho s}$.  The
minus sign in \eqref{eq:G_expanded} accounts for fermionic statistics:
if $\p{\mu,\nu}=\p{\sigma,\rho}$, then we are considering a term of
the form
\begin{align}
  G^{\mu q;\nu r}_{\nu s;\mu t}
  \c_{\nu s}^\dag \c_{\mu t}^\dag \c_{\nu r} \c_{\mu q}
  = -G^{qr}_{ts} \c_{\nu s}^\dag \c_{\mu t}^\dag \c_{\mu r} \c_{\nu q}
  = G^{qr}_{ts} \c_{\mu t}^\dag \c_{\nu s}^\dag \c_{\nu r} \c_{\mu q}.
\end{align}
At the cost of introducing an additional sum over new nuclear spin
indices, the expanded coupling tensor allows us to collect together
diagrams which have identical graph topologies, but address different
matrix elements of the effective Hamiltonian due to the exchange of
nuclear spins at a vertex.  The third order three-body diagrams in
$H_{\t{int}}^{(3)}$ are then
\begin{align}
  \begin{tikzpicture}
    \begin{feynman}
      \vertex (v1);
      \vertex[below right = of v1] (v2);
      \vertex[above right = of v2] (v3);
      \vertex[above left = of v1] (f1) {$\mu r$};
      \vertex[left = of v1] (f2) {$\nu s$};
      \vertex[below left = of v2] (f3) {$\rho t$};
      \vertex[above right = of v3] (f4) {$\mu r''$};
      \vertex[right = of v3] (f5) {$\nu's'''$};
      \vertex[below right = of v2] (f6) {$\rho't'$};
      \diagram* {
        (f1) -- (v1) --[scalar, edge label=$\ell\mu r'$] (v3) -- (f4),
        (f2) -- (v1)
        --[scalar, edge label'=$m\nu s'$] (v2)
        --[scalar, edge label'=$n\nu's''$] (v3)
        -- (f5),
        (f3) -- (v2) -- (f6), };
    \end{feynman}
  \end{tikzpicture}
  \propto K_{\ell m} K^m_n K_{\ell n}
  G^{rs}_{r's'} G^{\nu s';\rho t}_{\nu's'';\rho't'} G^{r's''}_{r''s'''}
  \c_{\mu r''}^\dag \c_{\nu's'''}^\dag \c_{\rho't'}^\dag
  \c_{\rho t} \c_{\nu s} \c_{\mu r},
  \label{eq:H_3_3_S}
\end{align}
\begin{align}
  \begin{tikzpicture}
    \begin{feynman}
      \vertex (v1);
      \vertex[right = of v1] (v2);
      \vertex[below right = of v2] (v3);
      \vertex[above left = of v1] (f1) {$\mu r$};
      \vertex[below left = of v1] (f2) {$\nu s$};
      \vertex[below left = of v3] (f3) {$\rho t$};
      \vertex[above right = of v2] (f4) {$\mu r''$};
      \vertex[above right = of v3] (f5) {$\nu s'''$};
      \vertex[below right = of v3] (f6) {$\rho t'$};
      \diagram* {
        (f1) -- (v1)
        --[half left, scalar, edge label=$\ell\mu r'$] (v2)
        -- (f4),
        (f2) -- (v1)
        --[half right, scalar, edge label'=$m\nu s'$] (v2)
        --[scalar, edge label=$n\nu s''$] (v3)
        -- (f5),
        (f3) -- (v3) -- (f6), };
    \end{feynman}
  \end{tikzpicture}
  \propto K_{\ell m} K^{\ell m}_n K_n
  G^{rs}_{r's'} G^{r's'}_{r''s''} G^{s''t}_{s'''t'}
  \c_{\mu r''}^\dag \c_{\nu s'''}^\dag \c_{\rho t'}^\dag
  \c_{\rho t} \c_{\nu s} \c_{\mu r},
  \label{eq:H_3_3_O}
\end{align}
and the mirror image of \eqref{eq:H_3_3_O}.  As prescribed by
$H_{\t{int}}^{(3)}$ in \eqref{eq:H_int_3}, these diagrams have an
associated negative sign if they contain only one excited virtual
state, and a factor of $1/2$ if they contain a virtual ground state.
Remembering that counter-terms are $\O\p{G^2}$, there are additionally
two third-order three-body diagrams in $H_{\t{int}}^{(2)}$, namely
\begin{align}
  \begin{tikzpicture}
    \begin{feynman}
      \vertex[crossed dot] (v1) {};
      \vertex[above left = 4em of v1] (f1) {$\mu r$};
      \vertex[below left = 4em of v1] (f2) {$\nu s$};
      \vertex[right = 4em of v1] (v2);
      \vertex[above right = 4em of v1] (f3) {$\mu r'$};
      \vertex[below left = of v2] (f4) {$\rho t$};
      \vertex[below right = of v2] (f5) {$\rho t'$};
      \vertex[above right = of v2] (f6) {$\nu s''$};
      \diagram* {
        (f1) -- (v1) -- (f3),
        (f2) -- (v1)
        --[scalar, edge label=$n\nu s'$] (v2),
        (f4) -- (v2) -- (f5),
        (v2) -- (f6), };
    \end{feynman}
  \end{tikzpicture}
  \propto K_n^2 \tilde G^{rs}_{r's'} G^{s't}_{s''t'}
  \c_{\mu r'}^\dag \c_{\nu s''}^\dag \c_{\rho t'}^\dag
  \c_{\rho t} \c_{\nu s} \c_{\mu r}
  \label{eq:H_3_3_counter_term}
\end{align}
and its mirror image, where $\tilde G^{qr}_{st}$ is equal to the
counter-term associated with $G^{qr}_{st}$.  To leading order in the
coupling constants, the renormalization condition in
\eqref{eq:counter_term} implies that
\begin{align}
  \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex[crossed dot] (v) {};
        \vertex[above left = of v] (f1);
        \vertex[below left = of v] (f2);
        \vertex[above right = of v] (f3);
        \vertex[below right = of v] (f4);
        \diagram* {
          (f1) -- (v),
          (f2) -- (v),
          (v) -- (f3),
          (v) -- (f4), };
      \end{feynman}
    \end{tikzpicture}}
  = \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[right = of v1] (v2);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[above right = of v2] (f3);
        \vertex[below right = of v2] (f4);
        \diagram* {
          (f1) -- (v1) --[half left, scalar] (v2) -- (f3),
          (f2) -- (v1) --[half right, scalar] (v2) -- (f4), };
      \end{feynman}
    \end{tikzpicture}},
\end{align}
which means that the sum over all counter-term diagrams in
\eqref{eq:H_3_3_counter_term} is
\begin{align}
  \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex[crossed dot] (v1) {};
        \vertex[right = 4em of v1] (v2);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[above right = of v1] (f3);
        \vertex[below left = of v2] (f4);
        \vertex[below right = of v2] (f5);
        \vertex[above right = of v2] (f6);
        \diagram* {
          (f1) -- (v1) -- (f3),
          (f2) -- (v1) --[scalar] (v2),
          (f4) -- (v2) -- (f5),
          (v2) -- (f6), };
      \end{feynman}
    \end{tikzpicture}}
  = \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[right = of v1] (v2);
        \vertex[below right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[below left = of v3] (f3);
        \vertex[above right = of v2] (f4);
        \vertex[above right = of v3] (f5);
        \vertex[below right = of v3] (f6);
        \diagram* {
          (f1) -- (v1) --[half left, scalar] (v2) -- (f4),
          (f2) -- (v1) --[half right, scalar] (v2)
          --[scalar] (v3) -- (f5),
          (f3) -- (v3) -- (f6), };
      \end{feynman}
    \end{tikzpicture}}.
  \label{eq:H_3_3_counter_term_sum}
\end{align}
The net contribution of the these counter-term diagrams together with
the two-particle-loop diagrams in \eqref{eq:H_3_3_O} is therefore
\begin{multline}
  - \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex[crossed dot] (v1) {};
        \vertex[right = 4em of v1] (v2);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[above right = of v1] (f3);
        \vertex[below left = of v2] (f4);
        \vertex[below right = of v2] (f5);
        \vertex[above right = of v2] (f6);
        \diagram* {
          (f1) -- (v1) -- (f3),
          (f2) -- (v1) --[scalar] (v2),
          (f4) -- (v2) -- (f5),
          (v2) -- (f6), };
      \end{feynman}
    \end{tikzpicture}}
  + \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[right = of v1] (v2);
        \vertex[below right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[below left = of v3] (f3);
        \vertex[above right = of v2] (f4);
        \vertex[above right = of v3] (f5);
        \vertex[below right = of v3] (f6);
        \diagram* {
          (f1) -- (v1) --[half left, scalar] (v2) -- (f4),
          (f2) -- (v1) --[half right, scalar] (v2)
          --[scalar] (v3) -- (f5),
          (f3) -- (v3) -- (f6), };
      \end{feynman}
    \end{tikzpicture}}
  - \f12 \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[right = of v1] (v2);
        \vertex[below right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[below left = of v3] (f3);
        \vertex[above right = of v2] (f4);
        \vertex[above right = of v3] (f5);
        \vertex[below right = of v3] (f6);
        \diagram* {
          (f1) -- (v1) --[half left, scalar] (v2) -- (f4),
          (f2) -- (v1) --[half right, scalar] (v2)
          -- (v3) -- (f5),
          (f3) -- (v3) -- (f6), };
      \end{feynman}
    \end{tikzpicture}}
  - \f12 \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[right = of v1] (v2);
        \vertex[below right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[below left = of v3] (f3);
        \vertex[above right = of v2] (f4);
        \vertex[above right = of v3] (f5);
        \vertex[below right = of v3] (f6);
        \diagram* {
          (f1) -- (v1) --[half left] (v2) -- (f4),
          (f2) -- (v1) --[half right] (v2)
          --[scalar] (v3) -- (f5),
          (f3) -- (v3) -- (f6), };
      \end{feynman}
    \end{tikzpicture}} \\
  = - \f12 \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[right = of v1] (v2);
        \vertex[below right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[below left = of v3] (f3);
        \vertex[above right = of v2] (f4);
        \vertex[above right = of v3] (f5);
        \vertex[below right = of v3] (f6);
        \diagram* {
          (f1) -- (v1) --[half left, scalar] (v2) -- (f4),
          (f2) -- (v1) --[half right, scalar] (v2)
          -- (v3) -- (f5),
          (f3) -- (v3) -- (f6), };
      \end{feynman}
    \end{tikzpicture}}
  - \f12 \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[right = of v1] (v2);
        \vertex[below right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[below left = of v3] (f3);
        \vertex[above right = of v2] (f4);
        \vertex[above right = of v3] (f5);
        \vertex[below right = of v3] (f6);
        \diagram* {
          (f1) -- (v1) --[half left] (v2) -- (f4),
          (f2) -- (v1) --[half right] (v2)
          --[scalar] (v3) -- (f5),
          (f3) -- (v3) -- (f6), };
      \end{feynman}
    \end{tikzpicture}}.
  \label{eq:H_3_3_cancellation}
\end{multline}
An identical cancellation occurs for the mirror images of the diagrams
in \eqref{eq:H_3_3_O} and \eqref{eq:H_3_3_counter_term}, which results
in a contribution to the effective Hamiltonian which is equal to
\eqref{eq:H_3_3_cancellation}.  The contribution from the
three-particle-loop diagrams in \eqref{eq:H_3_3_S} is
\begin{multline}
  \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[below right = of v1] (v2);
        \vertex[above right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[left = of v1] (f2);
        \vertex[below left = of v2] (f3);
        \vertex[above right = of v3] (f4);
        \vertex[right = of v3] (f5);
        \vertex[below right = of v2] (f6);
        \diagram* {
          (f1) -- (v1) --[scalar] (v3) -- (f4),
          (f2) -- (v1) --[scalar] (v2) --[scalar] (v3) -- (f5),
          (f3) -- (v2) -- (f6), };
      \end{feynman}
    \end{tikzpicture}}
  - \f12 \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[below right = of v1] (v2);
        \vertex[above right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[left = of v1] (f2);
        \vertex[below left = of v2] (f3);
        \vertex[above right = of v3] (f4);
        \vertex[right = of v3] (f5);
        \vertex[below right = of v2] (f6);
        \diagram* {
          (f1) -- (v1) -- (v3) -- (f4),
          (f2) -- (v1) --[scalar] (v2) -- (v3) -- (f5),
          (f3) -- (v2) -- (f6), };
      \end{feynman}
    \end{tikzpicture}}
  - \f12 \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[below right = of v1] (v2);
        \vertex[above right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[left = of v1] (f2);
        \vertex[below left = of v2] (f3);
        \vertex[above right = of v3] (f4);
        \vertex[right = of v3] (f5);
        \vertex[below right = of v2] (f6);
        \diagram* {
          (f1) -- (v1) -- (v3) -- (f4),
          (f2) -- (v1) -- (v2) --[scalar] (v3) -- (f5),
          (f3) -- (v2) -- (f6), };
      \end{feynman}
    \end{tikzpicture}} \\
  = \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[below right = of v1] (v2);
        \vertex[above right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[left = of v1] (f2);
        \vertex[below left = of v2] (f3);
        \vertex[above right = of v3] (f4);
        \vertex[right = of v3] (f5);
        \vertex[below right = of v2] (f6);
        \diagram* {
          (f1) -- (v1) --[scalar] (v3) -- (f4),
          (f2) -- (v1) --[scalar] (v2) --[scalar] (v3) -- (f5),
          (f3) -- (v2) -- (f6), };
      \end{feynman}
    \end{tikzpicture}}
  - \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[below right = of v1] (v2);
        \vertex[above right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[left = of v1] (f2);
        \vertex[below left = of v2] (f3);
        \vertex[above right = of v3] (f4);
        \vertex[right = of v3] (f5);
        \vertex[below right = of v2] (f6);
        \diagram* {
          (f1) -- (v1) -- (v3) -- (f4),
          (f2) -- (v1) --[scalar] (v2) -- (v3) -- (f5),
          (f3) -- (v2) -- (f6), };
      \end{feynman}
    \end{tikzpicture}}.
\end{multline}
The third-order three-body interaction Hamiltonian is therefore
\begin{align}
  H_3^{(3)} &= \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[below right = of v1] (v2);
        \vertex[above right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[left = of v1] (f2);
        \vertex[below left = of v2] (f3);
        \vertex[above right = of v3] (f4);
        \vertex[right = of v3] (f5);
        \vertex[below right = of v2] (f6);
        \diagram* {
          (f1) -- (v1) --[scalar] (v3) -- (f4),
          (f2) -- (v1) --[scalar] (v2) --[scalar] (v3) -- (f5),
          (f3) -- (v2) -- (f6), };
      \end{feynman}
    \end{tikzpicture}}
  - \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[below right = of v1] (v2);
        \vertex[above right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[left = of v1] (f2);
        \vertex[below left = of v2] (f3);
        \vertex[above right = of v3] (f4);
        \vertex[right = of v3] (f5);
        \vertex[below right = of v2] (f6);
        \diagram* {
          (f1) -- (v1) -- (v3) -- (f4),
          (f2) -- (v1) --[scalar] (v2) -- (v3) -- (f5),
          (f3) -- (v2) -- (f6), };
      \end{feynman}
    \end{tikzpicture}}
   - \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[right = of v1] (v2);
        \vertex[below right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[below left = of v3] (f3);
        \vertex[above right = of v2] (f4);
        \vertex[above right = of v3] (f5);
        \vertex[below right = of v3] (f6);
        \diagram* {
          (f1) -- (v1) --[half left, scalar] (v2) -- (f4),
          (f2) -- (v1) --[half right, scalar] (v2) -- (v3) -- (f5),
          (f3) -- (v3) -- (f6), };
      \end{feynman}
    \end{tikzpicture}}
  - \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[right = of v1] (v2);
        \vertex[below right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[below left = of v3] (f3);
        \vertex[above right = of v2] (f4);
        \vertex[above right = of v3] (f5);
        \vertex[below right = of v3] (f6);
        \diagram* {
          (f1) -- (v1) --[half left] (v2) -- (f4),
          (f2) -- (v1) --[half right] (v2) --[scalar] (v3) -- (f5),
          (f3) -- (v3) -- (f6), };
      \end{feynman}
    \end{tikzpicture}} \nonumber \\[1em]
  &= \p{\alpha_3^{(3)} - \alpha_5^{(3)}} \H_{3,1}^{(3)}
  - \p{\alpha_{4,3}^{(3)} + \alpha_5^{(3)}} \H_{3,2}^{(3)},
  \label{eq:H_3_3}
\end{align}
where, letting $E_{mn}\equiv E_m+E_n$,
\begin{align}
  \alpha_3^{(3)} \equiv \sum_{\substack{\ell+m>0\\\ell+n>0}}
  \f{K_{\ell m} K^m_n K_{\ell n}}{E_{\ell m} E_{\ell n}},
  &&
  \alpha_{4,3}^{(3)}
  \equiv K \sum_{m+n>0} \f{K_{mn}^2}{E_{mn}^2},
  &&
  \alpha_5^{(3)}
  \equiv  K \sum_{n>0} \f{K_n^2}{E_n^2},
\end{align}
\begin{align}
  \H_{3,1}^{(3)} \equiv \sum_{\abs{\set{\mu,\nu,\rho}}=3}
  G^{rs}_{r's'} G^{\nu s'\rho t}_{\nu's''\rho't'} G^{r's''}_{r''s'''}
  \c_{\mu r''}^\dag \c_{\nu's'''}^\dag \c_{\rho't'}^\dag
  \c_{\rho t} \c_{\nu s} \c_{\mu r}
\end{align}
\begin{align}
  \H_{3,2}^{(3)} \equiv \sum_{\abs{\set{\mu,\nu,\rho}}=3}
  G^{rs}_{r's'} G^{r's'}_{r''s''} G^{s''t}_{s'''t'}
  \c_{\mu r''}^\dag \c_{\nu s'''}^\dag \c_{\rho t'}^\dag
  \c_{\rho t} \c_{\nu s} \c_{\mu r}.
\end{align}

\subsection{Effective four-body interactions at third order}

At third order in the coupling constants, we have four-body terms of
the form
\begin{align}
  \begin{tikzpicture}
    \begin{feynman}
      \vertex[label=0:$~m\mu q'$] (v1);
      \vertex[below right = of v1, label=0:$~n\mu'q''$] (v2);
      \vertex[below right = of v2] (v3);
      \vertex[above left = of v1] (f1) {$\mu q$};
      \vertex[below left = of v1] (f2) {$\nu r$};
      \vertex[below left = of v2] (f3) {$\rho s$};
      \vertex[below left = of v3] (f4) {$\sigma t$};
      \vertex[above right = of v1] (f5) {$\nu r'$};
      \vertex[above right = of v2] (f6) {$\rho's'$};
      \vertex[above right = of v3] (f7) {$\sigma t'$};
      \vertex[below right = of v3] (f8) {$\mu'q'''$};
      \diagram* {
        (f1) -- (v1) -- (f5),
        (f2) -- (v1) -- (f5),
        (v1) --[scalar] (v2) --[scalar] (v3) -- (f8),
        (f3) -- (v2) -- (f6),
        (f4) -- (v3) -- (f7),
      };
    \end{feynman}
  \end{tikzpicture}
  \propto K_m K^m_n K_n
  G^{qr}_{q'r'} G^{\mu q'\rho s}_{\mu'q''\rho's'} G^{q''t}_{q'''t'}
  \c_{\mu'q'''}^\dag \c_{\nu r'}^\dag \c_{\rho's'}^\dag \c_{\sigma t'}^\dag
  \c_{\sigma t} \c_{\rho s} \c_{\nu r} \c_{\mu q},
  \label{eq:H_4_3_C}
\end{align}
\begin{align}
  \begin{tikzpicture}
    \begin{feynman}
      \vertex (v1);
      \vertex[above right = 1.5em of v1] (v2);
      \vertex[below right = 2.5em of v1] (v3);
      \vertex[right = 0.5em of v1] (label)
      {$\let\scriptstyle\textstyle\substack{m\mu q'\\n\nu r'}$};
      \vertex[above left = of v1] (f1) {$\mu q$};
      \vertex[below left = of v1] (f2) {$\nu r$};
      \vertex[above left = of v2] (f3) {$\rho s$};
      \vertex[below left = of v3] (f4) {$\sigma t$};
      \vertex[above right = of v2] (f5) {$\rho s'$};
      \vertex[right = of v2] (f6) {$\mu q''$};
      \vertex[right = of v3] (f7) {$\nu r''$};
      \vertex[below right = of v3] (f8) {$\sigma t'$};
      \diagram* {
        (f1) -- (v1) --[scalar] (v2) -- (f6),
        (f2) -- (v1) --[scalar] (v3) -- (f7),
        (f3) -- (v2) -- (f5),
        (f4) -- (v3) -- (f8),
      };
    \end{feynman}
  \end{tikzpicture}
  \propto K_{mn} K_m K_n G^{qr}_{q'r'} G^{q's}_{q''s'} G^{r't}_{r''t'}
  \c_{\mu q''}^\dag \c_{\nu r''}^\dag \c_{\rho s'}^\dag \c_{\sigma t'}^\dag
  \c_{\sigma t} \c_{\rho s} \c_{\nu r} \c_{\mu q},
  \label{eq:H_4_3_B}
\end{align}
and the mirror image of \eqref{eq:H_4_3_B}.  As we are computing the
leading-order contribution to effective four-body interactions, there
are no counter-terms contributions.  In principle, there is now also
the possibility to make the disconnected diagrams of the form
\begin{align}
  \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[right = of v1] (v2);
        \vertex[above right = of v2] (f3);
        \vertex[below right = of v2] (f4);
        \vertex[below = 1em of f2] (f5);
        \vertex[below = of f5] (f6);
        \vertex[below = 1em of f4] (f7);
        \vertex[below = of f7] (f8);
        \diagram* {
          (f1) -- (v1),
          (f2) -- (v1),
          (v2) -- (f3),
          (v2) -- (f4),
          (v1) --[scalar, half left] (v2),
          (v1) --[scalar, half right] (v2),
          (f5) -- (f8),
          (f6) -- (f7) };
      \end{feynman}
    \end{tikzpicture}},
  &&
  \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[right = of v1] (v2);
        \vertex[above right = of v2] (f3);
        \vertex[below right = of v2] (f4);
        \vertex[left = of f2] (a1);
        \vertex[left = of f4] (a2);
        \vertex[below = 1em of a1] (f5);
        \vertex[below = of f5] (f6);
        \vertex[below = 1em of a2] (f7);
        \vertex[below = of f7] (f8);
        \diagram* {
          (f1) -- (v1),
          (f2) -- (v1),
          (v2) -- (f3),
          (v2) -- (f4),
          (v1) --[scalar, half left] (v2),
          (v1) --[scalar, half right] (v2),
          (f5) -- (f8),
          (f6) -- (f7) };
      \end{feynman}
    \end{tikzpicture}},
  &&
  \t{and}
  &&
  \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[right = of v1] (v2);
        \vertex[above right = of v2] (f3);
        \vertex[below right = of v2] (f4);
        \vertex[right = of f2] (a1);
        \vertex[right = of f4] (a2);
        \vertex[below = 1em of a1] (f5);
        \vertex[below = of f5] (f6);
        \vertex[below = 1em of a2] (f7);
        \vertex[below = of f7] (f8);
        \diagram* {
          (f1) -- (v1),
          (f2) -- (v1),
          (v2) -- (f3),
          (v2) -- (f4),
          (v1) --[scalar, half left] (v2),
          (v1) --[scalar, half right] (v2),
          (f5) -- (f8),
          (f6) -- (f7) };
      \end{feynman}
    \end{tikzpicture}}.
\end{align}
As prescribed by $H_{\t{int}}^{(3)}$ in \eqref{eq:H_int_3}, however,
the second and third of these diagrams pick up a factor of $-1/2$, and
the sum of the disconnected diagrams vanishes.

Observing that the diagrams of the form in \eqref{eq:H_4_3_B} are
equal to their mirror image, the positive contributions to the
third-order four-body interaction Hamiltonian are
\begin{align}
  \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[above right = 1.5em of v1] (v2);
        \vertex[below right = 2.5em of v1] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[above left = of v2] (f3);
        \vertex[below left = of v3] (f4);
        \vertex[above right = of v2] (f5);
        \vertex[right = of v2] (f6);
        \vertex[right = of v3] (f7);
        \vertex[below right = of v3] (f8);
        \diagram* {
          (f1) -- (v1) --[scalar] (v2) -- (f6),
          (f2) -- (v1) --[scalar] (v3) -- (f7),
          (f3) -- (v2) -- (f5),
          (f4) -- (v3) -- (f8),
        };
      \end{feynman}
    \end{tikzpicture}}
  + \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[below right = 2.5em of v1] (v3);
        \vertex[below left = 1.7em of v3] (v2);
        \vertex[above left = of v1] (f1);
        \vertex[left = of v1] (f2);
        \vertex[left = of v2] (f3);
        \vertex[below left = of v2] (f4);
        \vertex[above right = of v1] (f5);
        \vertex[above right = of v3] (f6);
        \vertex[below right = of v3] (f7);
        \vertex[below right = of v2] (f8);
        \diagram* {
          (f1) -- (v1) -- (f5),
          (f2) -- (v1) --[scalar] (v3) -- (f6),
          (f3) -- (v2) --[scalar] (v3) -- (f7),
          (f4) -- (v2) -- (f8), };
      \end{feynman}
    \end{tikzpicture}}
  + \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[below right = of v1] (v2);
        \vertex[below right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[below left = of v2] (f3);
        \vertex[below left = of v3] (f4);
        \vertex[above right = of v1] (f5);
        \vertex[above right = of v2] (f6);
        \vertex[above right = of v3] (f7);
        \vertex[below right = of v3] (f8);
        \diagram* {
          (f1) -- (v1) -- (f5),
          (f2) -- (v1) -- (f5),
          (v1) --[scalar] (v2) --[scalar] (v3) -- (f8),
          (f3) -- (v2) -- (f6),
          (f4) -- (v3) -- (f7),
        };
      \end{feynman}
    \end{tikzpicture}}
  = 2 \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[above right = 1.5em of v1] (v2);
        \vertex[below right = 2.5em of v1] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[above left = of v2] (f3);
        \vertex[below left = of v3] (f4);
        \vertex[above right = of v2] (f5);
        \vertex[right = of v2] (f6);
        \vertex[right = of v3] (f7);
        \vertex[below right = of v3] (f8);
        \diagram* {
          (f1) -- (v1) --[scalar] (v2) -- (f6),
          (f2) -- (v1) --[scalar] (v3) -- (f7),
          (f3) -- (v2) -- (f5),
          (f4) -- (v3) -- (f8),
        };
      \end{feynman}
    \end{tikzpicture}}
  + \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[below right = of v1] (v2);
        \vertex[below right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[below left = of v2] (f3);
        \vertex[below left = of v3] (f4);
        \vertex[above right = of v1] (f5);
        \vertex[above right = of v2] (f6);
        \vertex[above right = of v3] (f7);
        \vertex[below right = of v3] (f8);
        \diagram* {
          (f1) -- (v1) -- (f5),
          (f2) -- (v1) -- (f5),
          (v1) --[scalar] (v2) --[scalar] (v3) -- (f8),
          (f3) -- (v2) -- (f6),
          (f4) -- (v3) -- (f7),
        };
      \end{feynman}
    \end{tikzpicture}},
\end{align}
while the negative contributions are
\begin{multline}
  \f12 \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[above right = 1.5em of v1] (v2);
        \vertex[below right = 2.5em of v1] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[above left = of v2] (f3);
        \vertex[below left = of v3] (f4);
        \vertex[above right = of v2] (f5);
        \vertex[right = of v2] (f6);
        \vertex[right = of v3] (f7);
        \vertex[below right = of v3] (f8);
        \diagram* {
          (f1) -- (v1) --[scalar] (v2) -- (f6),
          (f2) -- (v1) -- (v3) -- (f7),
          (f3) -- (v2) -- (f5),
          (f4) -- (v3) -- (f8),
        };
      \end{feynman}
    \end{tikzpicture}}
  + \f12 \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[below right = 2.5em of v1] (v3);
        \vertex[below left = 1.7em of v3] (v2);
        \vertex[above left = of v1] (f1);
        \vertex[left = of v1] (f2);
        \vertex[left = of v2] (f3);
        \vertex[below left = of v2] (f4);
        \vertex[above right = of v1] (f5);
        \vertex[above right = of v3] (f6);
        \vertex[below right = of v3] (f7);
        \vertex[below right = of v2] (f8);
        \diagram* {
          (f1) -- (v1) -- (f5),
          (f2) -- (v1) -- (v3) -- (f6),
          (f3) -- (v2) --[scalar] (v3) -- (f7),
          (f4) -- (v2) -- (f8), };
      \end{feynman}
    \end{tikzpicture}}
  + \f12 \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[below right = of v1] (v2);
        \vertex[below right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[below left = of v2] (f3);
        \vertex[below left = of v3] (f4);
        \vertex[above right = of v1] (f5);
        \vertex[above right = of v2] (f6);
        \vertex[above right = of v3] (f7);
        \vertex[below right = of v3] (f8);
        \diagram* {
          (f1) -- (v1) -- (f5),
          (f2) -- (v1) -- (f5),
          (v1) --[scalar] (v2) -- (v3) -- (f8),
          (f3) -- (v2) -- (f6),
          (f4) -- (v3) -- (f7),
        };
      \end{feynman}
    \end{tikzpicture}}
  + \f12 \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[below right = of v1] (v2);
        \vertex[below right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[below left = of v2] (f3);
        \vertex[below left = of v3] (f4);
        \vertex[above right = of v1] (f5);
        \vertex[above right = of v2] (f6);
        \vertex[above right = of v3] (f7);
        \vertex[below right = of v3] (f8);
        \diagram* {
          (f1) -- (v1) -- (f5),
          (f2) -- (v1) -- (f5),
          (v1) -- (v2) --[scalar] (v3) -- (f8),
          (f3) -- (v2) -- (f6),
          (f4) -- (v3) -- (f7),
        };
      \end{feynman}
    \end{tikzpicture}} \\
  = \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[above right = 1.5em of v1] (v2);
        \vertex[below right = 2.5em of v1] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[above left = of v2] (f3);
        \vertex[below left = of v3] (f4);
        \vertex[above right = of v2] (f5);
        \vertex[right = of v2] (f6);
        \vertex[right = of v3] (f7);
        \vertex[below right = of v3] (f8);
        \diagram* {
          (f1) -- (v1) --[scalar] (v2) -- (f6),
          (f2) -- (v1) -- (v3) -- (f7),
          (f3) -- (v2) -- (f5),
          (f4) -- (v3) -- (f8),
        };
      \end{feynman}
    \end{tikzpicture}}
  + \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[below right = of v1] (v2);
        \vertex[below right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[below left = of v2] (f3);
        \vertex[below left = of v3] (f4);
        \vertex[above right = of v1] (f5);
        \vertex[above right = of v2] (f6);
        \vertex[above right = of v3] (f7);
        \vertex[below right = of v3] (f8);
        \diagram* {
          (f1) -- (v1) -- (f5),
          (f2) -- (v1) -- (f5),
          (v1) --[scalar] (v2) -- (v3) -- (f8),
          (f3) -- (v2) -- (f6),
          (f4) -- (v3) -- (f7),
        };
      \end{feynman}
    \end{tikzpicture}},
\end{multline}
such that the net third-order four-body interaction Hamiltonian is
\begin{align}
  H_4^{(3)} &= 2 \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[above right = 1.5em of v1] (v2);
        \vertex[below right = 2.5em of v1] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[above left = of v2] (f3);
        \vertex[below left = of v3] (f4);
        \vertex[above right = of v2] (f5);
        \vertex[right = of v2] (f6);
        \vertex[right = of v3] (f7);
        \vertex[below right = of v3] (f8);
        \diagram* {
          (f1) -- (v1) --[scalar] (v2) -- (f6),
          (f2) -- (v1) --[scalar] (v3) -- (f7),
          (f3) -- (v2) -- (f5),
          (f4) -- (v3) -- (f8),
        };
      \end{feynman}
    \end{tikzpicture}}
  - \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[above right = 1.5em of v1] (v2);
        \vertex[below right = 2.5em of v1] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[above left = of v2] (f3);
        \vertex[below left = of v3] (f4);
        \vertex[above right = of v2] (f5);
        \vertex[right = of v2] (f6);
        \vertex[right = of v3] (f7);
        \vertex[below right = of v3] (f8);
        \diagram* {
          (f1) -- (v1) --[scalar] (v2) -- (f6),
          (f2) -- (v1) -- (v3) -- (f7),
          (f3) -- (v2) -- (f5),
          (f4) -- (v3) -- (f8),
        };
      \end{feynman}
    \end{tikzpicture}}
  + \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[below right = of v1] (v2);
        \vertex[below right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[below left = of v2] (f3);
        \vertex[below left = of v3] (f4);
        \vertex[above right = of v1] (f5);
        \vertex[above right = of v2] (f6);
        \vertex[above right = of v3] (f7);
        \vertex[below right = of v3] (f8);
        \diagram* {
          (f1) -- (v1) -- (f5),
          (f2) -- (v1) -- (f5),
          (v1) --[scalar] (v2) --[scalar] (v3) -- (f8),
          (f3) -- (v2) -- (f6),
          (f4) -- (v3) -- (f7),
        };
      \end{feynman}
    \end{tikzpicture}}
  - \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[below right = of v1] (v2);
        \vertex[below right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[below left = of v2] (f3);
        \vertex[below left = of v3] (f4);
        \vertex[above right = of v1] (f5);
        \vertex[above right = of v2] (f6);
        \vertex[above right = of v3] (f7);
        \vertex[below right = of v3] (f8);
        \diagram* {
          (f1) -- (v1) -- (f5),
          (f2) -- (v1) -- (f5),
          (v1) --[scalar] (v2) -- (v3) -- (f8),
          (f3) -- (v2) -- (f6),
          (f4) -- (v3) -- (f7),
        };
      \end{feynman}
    \end{tikzpicture}} \nonumber \\
  &= \p{2\alpha_{4,1}^{(3)} - \alpha_5^{(3)}} \H_{4,1}^{(3)}
  + \p{\alpha_{4,2}^{(3)} - \alpha_5^{(3)}} \H_{4,2}^{(3)},
  \label{eq:H_4_3}
\end{align}
where
\begin{align}
  \alpha_{4,1}^{(3)}
  \equiv \sum_{\substack{m\ge0\\n>0}} \f{K_{mn} K_m K_n}{E_{mn} E_n},
  &&
  \alpha_{4,2}^{(3)}
  \equiv \sum_{m,n>0} \f{K_m K^m_n K_n}{E_m E_n},
  &&
  \alpha_5^{(3)}
  \equiv  K \sum_{n>0} \f{K_n^2}{E_n^2},
\end{align}
\begin{align}
  \H_{4,1}^{(3)}
  \equiv \sum_{\abs{\set{\mu,\nu,\rho,\sigma}}=4}
  G^{qr}_{q'r'} G^{q's}_{q''s'} G^{r't}_{r''t'}
  \c_{\mu q''}^\dag \c_{\nu r''}^\dag \c_{\rho s'}^\dag \c_{\sigma t'}^\dag
  \c_{\sigma t} \c_{\rho s} \c_{\nu r} \c_{\mu q},
\end{align}
\begin{align}
  \H_{4,2}^{(3)}
  \equiv \sum_{\abs{\set{\mu,\nu,\rho,\sigma}}=4}
  G^{qr}_{q'r'} G^{\mu q'\rho s}_{\mu'q''\rho's'} G^{q''t}_{q'''t'}
  \c_{\mu'q'''}^\dag \c_{\nu r'}^\dag \c_{\rho's'}^\dag \c_{\sigma t'}^\dag
  \c_{\sigma t} \c_{\rho s} \c_{\nu r} \c_{\mu q}.
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Low-excitation Hamiltonians, eigenstates, and spectra}
\label{sec:spectra}

Current experiments with ultracold ${}^{87}$Sr on a lattice can
coherently address ground states and their low-lying orbital
excitations for up to five atoms per lattice site
\cite{goban2018emergence}.  Due to the SU($N$) symmetry of
inter-atomic interactions, manifest in the fact that all coupling
constants are independent of nuclear spin, we can write all effective
$M$-body Hamiltonians which address these states in the form
\begin{align}
  H_M = \sum_{\abs{\set{\mu_j}}=M}
  \p{U_{M,\g} \prod_{\alpha=1}^M \n_{\mu_\alpha,\g}
    + U_{M,+} \n_{\mu_1,\e} \prod_{\alpha=2}^M \n_{\mu_\alpha,\g}
    + U_{M,-} \c_{\mu_1,\g}^\dag \c_{\mu_2,\e}^\dag
    \c_{\mu_2,\g} \c_{\mu_1,\e} \prod_{\alpha=3}^M \n_{\mu_\alpha,\g}},
  \label{eq:H_M}
\end{align}
where $\n_{\mu s}\equiv \c_{\mu s}^\dag\c_{\mu s}$ is a number
operator, and the coefficients $U_X$ can be determined from the
coupling constants $G_Y$ and prefactors $\alpha_Z^{(p)}$ of the
effective $M$-body Hamiltonians derived in section
\ref{sec:low_energy} (see Appendix \ref{sec:U_X}).  For a lattice with
$N\ge M$ atoms occupying nuclear spins $\N=\set{\mu_j}$ for
$j=1,\cdots,N$, the $M$-body Hamiltonian $H_M$ has a single ground
state $\ket{\N,0}$, and a singly-excited state $\ket{\N,+}$ which is
fully symmetric in the orbital degrees of freedom; these states are
\begin{align}
  \ket{\N,0}
  \equiv \p{\prod_{\mu\in\N} \c_{\mu,\g}^\dag} \ket{\t{vacuum}},
  &&
  \ket{\N,+} \equiv \f1{\sqrt{N}} \sum_{\mu\in\N}
  \c_{\mu,\e}^\dag \c_{\mu,\g} \ket{\N,0}.
  \label{eq:states_S}
\end{align}
As these states are fully symmetric in their orbital degrees of
freedom, they are anti-symmetric in their nuclear spin degrees of
freedom, forming an SU($N$) singlet.  Furthermore, the symmetric state
is particularly interesting as its orbital degrees of freedom form an
$N$-body entangled $W$ state, which belongs to a special class of
multi-partite entangled states that are robust against disposal or
loss of particles.  This state thus constitutes an important resource
for many quantum information processing and quantum communication
tasks \cite{zang2015generating}.

In addition to the states in \eqref{eq:states_S}, the multi-body
Hamiltonian $H_M$ in \eqref{eq:H_M} has an $\p{N-1}$-fold degenerate
excited-state eigenspace which is asymmetric in the orbital degrees of
freedom, spanned by the states
\begin{align}
  \ket{\N,-_j} \equiv \f1{\sqrt2}
  \p{\c_{\mu_1,\e}^\dag \c_{\mu_1,\g} - \c_{\mu_j,\e}^\dag \c_{\mu_j,\g}}
  \ket{\N,0}
  \label{eq:states_A}
\end{align}
for $j=2,\cdots,N$.  If $N>2$, the asymmetric states are not separable
in their orbital and nuclear spin degrees of freedom.  An important
feature of the excited states in \eqref{eq:states_S} and
\eqref{eq:states_A} is that they are entirely independent of $M$,
which implies that the effect of multi-body interactions is simply to
modify the many-body atomic energy spectra without affecting the
energy eigenstates.  The eigenvalues $E_{NX}^{(M)}=\bk{\N X|H_M|\N X}$
of $H_M$ associated with each of the eigenstates in
\eqref{eq:states_S} and \eqref{eq:states_A} are provided in table
\ref{tab:eigen}, both in terms of the coefficients $U_{MX}$ of $H_M$
as expressed in \eqref{eq:H_M} and the $M$-body eigenvalues
$E_{MX}^{(M)}$.  Due to the SU($N$) symmetry of the multi-body
Hamiltonian $H_M$, the eigenvalues $E_{NX}^{(M)}$ depend on the number
of nuclear spins on a lattice site, $N$, but not on the actual nuclear
spins $\mu\in\N$ which are occupied.  The total $N$-body interaction
energies $E_{NX}$ are given in terms of the $M$-body eigenvalues
$E_{NX}^{(M)}$ by $E_{NX}=\sum_ME_{NX}^{(M)}$.

\begin{table}
  \centering
  \caption{\footnotesize Low-excitation eigenvalues of $M$-body
    Hamiltonians $H_M$.  Many-body energy eigenstates are labeled by
    the nuclear spins they occupy (i.e.~$\N$ with $\abs{\N}=N$) and
    whether they are in the orbital ground ($0$), singly-excited
    symmetric ($+$), or singly-excited asymmetric ($-$) state.  The
    corresponding $N$-body eigenvalues $E_{NX}^{(M)}$ of $H_M$ are
    given in terms of the coefficients $U_{MX}$ as appearing in
    \eqref{eq:H_M}, in addition to the $M$-body eigenvalues
    $E_{MX}^{(M)}$.}
  \label{tab:eigen}
  \begin{tabular}{c|c}
    Eigenstate
    & $H_M$ eigenvalue \\ \hline\hline
    $\ket{\N,0}$
    & $M! {N \choose M} U_{M,\g}$ \\ \hline
    $\ket{\N,+}$
    & $M! {N-1 \choose M} U_{M,\g}
      + \p{M-1}! {N-1 \choose M-1} \p{U_{M,+} + U_{M,-}}$ \\ \hline
    $\ket{\N,-}$
    & $M! {N-1 \choose M} U_{M,\g}
      + \p{M-1}! {N-1 \choose M-1} U_{M,+}
      - \p{M-2}! {N-2 \choose M-2} U_{M,-}$ \\ \hline\hline
    $\ket{\N,0}$
    & ${N \choose M} E_{M,0}^{(M)}$ \\ \hline
    $\ket{\N,+}$
    & ${N-1 \choose M} E_{M,0}^{(M)}
      + {N-1 \choose M-1} E_{M,+}^{(M)}$ \\ \hline
    $\ket{\N,-}$
    & ${N-1 \choose M} E_{M,0}^{(M)}
      + M^{-1} \sp{{N-1 \choose M-1} - {N-2 \choose M-2}} E_{M,+}^{(M)}
      + \sp{\p{1 - M^{-1}} {N-1 \choose M-1}
      + M^{-1} {N-2 \choose M-2}} E_{M,-}^{(M)}$
  \end{tabular}
\end{table}

\subsection{Many-body state spectroscopy}

Spectroscopic interrogation is a powerful means to probe the internal
structure and dynamics of a system under examination.  Consequently,
we consider Rabi spectroscopy of the low-lying energy eigenstates in
multiply-occupied lattice sites.  If we interrogate a lattice site by
a laser red-detuned by $\Delta$ from the single-atom orbital state
excitation energy, we realize the Hamiltonian
\begin{align}
  H_{\t{Rabi}}
  = \sum_X E_X \P_X
  + \sum_\mu\p{\Delta T_\mu^z + \Omega_\mu T_\mu^x},
  \label{eq:H_rabi}
\end{align}
where $E_X$ is an eigenvalue of the effective interaction Hamiltonian
$H_{\t{int}}^{\t{eff}}$, $\P_X$ is a projector onto the corresponding
eigenspace, and
\begin{align}
  T_\mu^z \equiv \f12 \p{\c_{\mu,\e}^\dag \c_{\mu,\e}
    - \c_{\mu,\g}^\dag \c_{\mu,\g}},
  &&
  T_\mu^x \equiv \f12 \p{\c_{\mu,\e}^\dag \c_{\mu,\g}
    + \c_{\mu,\g}^\dag \c_{\mu,\e}},
\end{align}
are single-atom pseudospin operators.  The Rabi frequency $\Omega_\mu$
is proportional to the Clebsch-Gordan coefficient
$\bk{I,\mu;1,0|I,\mu}\sim\mu$ for a photon-induced
nuclear-spin-conserving orbital state transition of an atom with
nuclear spin $\mu$.  We therefore define the bare Rabi frequency
$\Omega_0\equiv\Omega_\mu/\mu$ to explicitly factor out dependence on
nuclear spins $\mu$.

Consider now a single lattice site in the orbital ground state
$\ket{\N,0}$ with nuclear spins $\N\equiv\set{\mu_j}$ for
$j=1,\cdots,N$.  If we red-detune the interrogation laser by $\delta$
from a many-body orbital state excitation energy, i.e.~set
$\Delta=\Delta_{NX}-\delta$ for $\Delta_{NX}\equiv E_{NX}-E_{N,0}$ and
$X\in\set{+,-}$, then in the subspace of the target states
$\set{\ket{\N,0},\ket{\N X}}$ the Hamiltonian in \eqref{eq:H_rabi}
becomes
\begin{align}
  H_{\N X} = \delta S_{\N X}^z + \Omega_{\N X} S_{\N X}^x,
  \label{eq:H_NX}
\end{align}
where
\begin{align}
  S_{\N X}^z &\equiv \f12\p{\op{\N X}-\op{\N,0}},
  \label{eq:S_NX_z} \\
  S_{\N X}^x &\equiv \f12\p{\op{\N X}{\N,0}+\op{\N,0}{\N X}},
  \label{eq:S_NX_x}
\end{align}
are many-body pseudospin operators, and the Rabi frequencies
$\Omega_{\N X}$ are determined by
\begin{align}
  H_{\t{Rabi}}\ket{\N,0}
  = \f12\Omega_0\sum_\mu\mu\c_{\mu,\e}^\dag\c_{\mu,\g}\ket{\N,0}
  = \f12\Omega_{\N,+}\ket{\N,+} + \f12\Omega_{\N,-}\ket{\N,-}.
  \label{eq:O_NX}
\end{align}
While the symmetric excited state $\ket{\N,+}$ is given in
\eqref{eq:states_S}, at this point we have not explicitly solved for
the asymmetric excited state $\ket{\N,-}$.  The asymmetric state is
implicitly defined by \eqref{eq:O_NX}, and lies somewhere in the span
of the $N-1$ asymmetric states given in \eqref{eq:states_A}.
Determining the symmetric-state Rabi frequency $\Omega_{\N,+}$ is
simply a matter of projecting the expression in \eqref{eq:O_NX} onto
$\ket{\N,+}$, which yields
\begin{align}
  \Omega_{\N,+}
  = \bk{\N,+|\Omega_0\sum_\mu\mu\c_{\mu,\e}^\dag\c_{\mu,\g}|\N,0}
  = \Omega_0 \sum_{\mu\in\N} \f{\mu}{\sqrt{N}}
  = \Omega_0 \sqrt{N} \bar\mu_\N,
  \label{eq:O_N+}
\end{align}
where $\bar\mu_\N\equiv\sum_{\mu\in\N}\mu/N$ is the average nuclear
spin $\mu\in\N$.  In order to determine the asymmetric-state Rabi
frequency $\Omega_{\N,-}$, we rearrange \eqref{eq:O_NX} to find
\begin{align}
  \Omega_{\N,-}\ket{\N,-}
  = \Omega_0\sum_\mu\mu\c_{\mu,\e}^\dag\c_{\mu,\g}\ket{\N,0}
  - \Omega_{\N,+}\ket{\N,+}
  = \Omega_0 \sum_{\mu\in\N} \p{\mu - \bar\mu_\N}
  \c_{\mu,\e}^\dag \c_{\mu,\g} \ket{\N,0}.
\end{align}
Denoting the standard deviation of nuclear spins $\mu\in\N$ by
$\sigma_\N$, normalization of $\ket{\N,-}$ thus determines the
asymmetric-state Rabi frequency
\begin{align}
  \Omega_{\N,-}
  = \Omega_0 \sp{\sum_{\mu\in\N}\p{\mu-\bar\mu_\N}^2}^{1/2}
  = \Omega_0 \sqrt{N} \sigma_\N,
  \label{eq:O_N-}
\end{align}
which in turn implies that the asymmetric excited state $\ket{\N,-}$
is
\begin{align}
  \ket{\N,-} = \f1{\sqrt{N}}\sum_{\mu\in\N}
  \p{\f{\mu-\bar\mu_\N}{\sigma_\N}}
  \c_{\mu,\e}^\dag \c_{\mu,\g} \ket{\N,0}.
\end{align}

\begin{figure}
  \centering
  \includegraphics{coefficients_table.pdf}
  \caption{\footnotesize Multiplicities of the magnitudes of reduced
    Rabi frequencies
    $\omega_{\N X}\equiv\Omega_{\N X}/\Omega_0\sqrt{N}$ in a lattice
    with a uniform mixture of nuclear spins with $I=9/2$ and
    single-site occupation numbers $N$ which are achievable in current
    ${}^{87}$Sr experiments.}
  \label{fig:coefficients}
\end{figure}

Figure \ref{fig:coefficients} shows multiplicities of the magnitudes
of reduced Rabi frequencies
$\omega_{\N X}\equiv\Omega_{\N X}/\Omega_0\sqrt{N}$ in a lattice with
a uniform mixture of nuclear spins with $I=9/2$ for single-site
occupation numbers $N$ which are achievable in current ${}^{87}$Sr
experiments \cite{goban2018emergence}.  On average, asymmetric-state
Rabi frequencies are greater in magnitude, which becomes more
pronounced for larger single-site occupation numbers.

\subsection{Experimental signatures and comparison}

\begin{figure}
  \centering
  \subfloat{\includegraphics{sweep_V54_O50.pdf}}
  \subfloat{\includegraphics{sweep_expt_V54.pdf}}
  \caption{\footnotesize Population (in arbitrary units) of the
    excited ${}^3P_0$ orbital state of ${}^{87}$Sr atoms in a uniform
    mixture of nuclear spins.  Atoms are prepared in the ground state
    of a lattice with depth $\U=54E_R$, where
    $E_R\approx3.5\times2\pi~\t{kHz}$ is the lattice photon recoil
    energy of the atoms, and interrogated by a laser with Rabi
    frequency $\Omega_I=50\times2\pi~\t{Hz}$ for a time
    $t=\pi/\Omega_I$.  ({\bf Left}) Populations predicted by the
    low-energy effective theory (with $s$-wave scattering parameters
    retrieved from ref.~\cite{zhang2014spectroscopic}), averaged over
    all nuclear spin combinations of $N\in\set{1,\cdots,5}$ atoms per
    lattice site for a fixed total atom number.  ({\bf Right})
    Experimental measurements of ${}^3P_0$ populations retrieved from
    ref.~\cite{goban2018emergence}, with Lorentzian fits to each peak
    as a visual guide.  Resonance peaks are identified by the
    many-body orbital states which are excited at the peak.}
  \label{fig:sweep}
\end{figure}

We now consider samples of ${}^{87}$Sr atoms in a uniform mixture of
nuclear spins $\mu\in\set{-I,\cdots,I}$ prepared in motional ground
states of a rectangular lattice with depths
$\U=\p{\U_x,\U_y,\U_z}=\p{41,55,69}E_R$, where
$E_R\approx3.5\times2\pi~\t{kHz}$ is the lattice photon recoil energy
of the atoms.  Such samples can be prepared in experiments which can
vary the single-site occupation number $N$, and which can control for
the total number of atoms that are addressed by an external
interrogation laser.  Figure \ref{fig:sweep} shows the population of
the excited ${}^3P_0$ orbital state when these atoms are interrogated
for a time $t=\pi/\Omega_I$ by a laser with Rabi frequency
$\Omega_I=50\times2\pi~\t{Hz}$ (i.e.~for individual atoms with nuclear
spin $\mu=I$) and detuning $\Delta$ from the single-atom
${}^1S_0\to{}^3P_0$ orbital excitation energy.  The ${}^3P_0$
population peaks when the laser detuning $\Delta$ is equal to the
many-body excitation energy $\Delta_{NX}\equiv E_{NX}-E_{N,0}$ for
$X\in\set{+,-}$, as this is precisely when the on-resonance condition
$\delta=0$ is satisfied in the many-body Rabi Hamiltonian $H_{\N X}$
in \eqref{eq:H_NX}.  Due to experimental uncertainties which vary with
single-site occupation number $N$, the heights of experimental peaks
in figure \ref{fig:sweep} are not well-calibrated between different
values of $N$.  Nonetheless, figure \ref{fig:sweep} exhibits
signatures of larger asymmetric-state Rabi frequencies than
symmetric-state ones in the form of higher asymmetric-state peaks for
fixed $N$.

\begin{figure}
  \centering
  \includegraphics{shifts_table.pdf}
  \caption{\footnotesize Multi-body excitation energies of ultracold
    ${}^{87}$Sr atoms at various lattice depths.  The top plot in each
    sub-figure with fixed $N,X$ shows the excitation energies
    $\Delta_{NX}\equiv E_{NX}-E_{N,0}$ measured experimentally in
    ref.~\cite{goban2018emergence} and those predicted by the
    low-energy effective theory at different orders in the coupling
    constants, when applicable both with and without four-body
    contributions.  The bottom plot in each sub-figure shows the
    relative error
    $\eta_{NX}\equiv
    \Delta_{NX}^{\t{theory}}/\Delta_{NX}^{\t{experiment}}-1$.  Error
    bars represent experimental error or conservatively estimated
    theoretical uncertainties from nearest-neighbor hopping of virtual
    states in the low-energy effective theory (see Appendix
    \ref{sec:error}).}
  \label{fig:shifts}
\end{figure}

\begin{figure}
  \centering
  \includegraphics{shifts_summary.pdf}
  \caption{\footnotesize Summary of the many-body excitation spectra
    in figure \ref{fig:shifts}, retrieved from
    ref.~\cite{goban2018emergence}.}
  \label{fig:shifts_summary}
\end{figure}

Identifying peaks in excitation spectra such as in figure
\ref{fig:sweep} constitutes a measurement of many-body excitation
energies, which was performed in ref.~\cite{goban2018emergence} to
detect signatures of effective multi-body interactions.  Figure
\ref{fig:shifts} shows a comparison between experimental measurements
of the many-body excitation energies $\Delta_{NX}$ for
$N\in\set{3,4,5}$ at various lattice depths $\U$, and the
corresponding values of $\Delta_{NX}$ predicted by the low-energy
effective theory at different orders in the coupling constants.  A
summary of the data in figure \ref{fig:shifts} is provided in figure
\ref{fig:shifts_summary}.  These figures show a clear improvement of
the low-energy effective theory we have developed with increasing
order in the coupling constants, in particular demonstrating the need
to consider multi-body interactions in a single-band model of the
atoms without motional excitations.

We expect that the dominant source of error in the third-order
effective theory comes from the fact that we neglect the inter-site
matrix elements of all Hamiltonians.  This source of error is
discussed in Appendix \ref{sec:error}, and leads to theoretical
uncertainties represented by error bars in figure \ref{fig:shifts}.
As these uncertainties are sometimes unable to account for the
residual disagreement between experimental measurements and the
effective theory, better agreement requires either a better
understanding of experimental procedures and error, a better
understanding of theoretical uncertainties, or consideration of yet
higher-order terms in the effective theory.  Due to a combinatorial
explosion in the number of diagrams which appear at increasing
perturbative orders of the effective theory, we need more systematic
methods to compute effective multi-body Hamiltonians at fourth order;
we therefore leave this calculation for future work.

\subsection{Orbital-state dynamics of a nuclear spin mixture}

In addition to spectral measurements of many-body interaction
energies, we consider the dynamics of multiply-occupied lattice sites
during spectroscopic interrogation.  While these dynamics do not
provide information about the nature or origin of effective multi-body
interactions, they provide tools and intuition for addressing the
low-lying orbital excitations which are readily accessible in an
experimental setting.  If we initialize all atoms in the $N$-body
ground state with an incoherent mixture of all nuclear spins, then we
prepare the mixed state $\rho_{N,0}=\P_{N,0}/\Tr\P_{N,0}$, where
\begin{align}
  \P_{NX} \equiv \sum_{\abs{\N}=N}\op{\N X}
  \label{eq:P_NX}
\end{align}
is a projector onto the space of the $N$-body orbital states
$\ket{\N X}$.  Interrogating the atoms for a time $t$ by a laser
resonant with the excitation energy $\Delta_{N,\pm}$ then gives us the
state
\begin{align}
  \rho_N^{(\pm)}\p{t} = \f1{\Tr\P_{N,0}}
  \sum_{\abs{\N}=N} \exp\p{-it\Omega_{\N,\pm}S_{\N,\pm}^x}
  \op{\N,0} \exp\p{it\Omega_{\N,\pm}S_{\N,\pm}^x},
\end{align}
where the Rabi frequencies $\Omega_{\N,+},\Omega_{\N,-}$ and
pseudo-spin operators $S_{\N,\pm}^x$ are respectively given in
\eqref{eq:O_N+}, \eqref{eq:O_N-}, and \eqref{eq:S_NX_x}.  Denoting the
eigenstates of $S_{\N,\pm}^x$ by
\begin{align}
  \ket{\N,\S_\pm} \equiv \f1{\sqrt2} \p{\ket{\N,0} + \ket{\N,\pm}},
  &&
  \ket{\N,\A_\pm} \equiv \f1{\sqrt2} \p{\ket{\N,0} - \ket{\N,\pm}},
\end{align}
and defining the identity operator projected to the relevant subspace,
\begin{align}
  \1_{\N,\pm} \equiv \op{\N,0} + \op{\N,\pm}
  = \op{\N,\S_\pm} + \op{\N,\A_\pm},
\end{align}
we can write the state $\rho_N^{(\pm)}\p{t}$ and excited-state
projectors $\P_{N,\pm}$ in the form
\begin{align}
  \rho_N^{(\pm)}\p{t} = \f1{\Tr\P_{N,0}} \sum_{\abs{\N}=N} \f12
  \sp{\1_{\N,\pm} + e^{i2t\Omega_{\N,\pm}}\op{\N,\S_\pm}{\N,\A_\pm}
    + e^{-i2t\Omega_{\N,\pm}}\op{\N,\A_\pm}{\N,\S_\pm}},
  \label{eq:rho_N_pm_t}
\end{align}
and
\begin{align}
  \P_{N,\pm} = \sum_{\abs{\N}=N} \f12
  \sp{\1_{\N,\pm}-\op{\N,\S_\pm}{\N,\A_\pm}-\op{\N,\A_\pm}{\N,\S_\pm}},
\end{align}
from which it follows that the net excited-state population at time
$t$ is
\begin{align}
  \bk{\P_{N,\pm}\p{t}}
  \equiv \Tr\sp{\rho_N^{(\pm)}\p{t}\P_{N,\pm}}
  = \f12 - \f12 \bk{\cos\p{2t\Omega_{\N,\pm}}}_{\abs{\N}=N},
  \label{eq:P_N_pm_t}
\end{align}
where $\bk{X}_{\abs{\N}=N}\equiv\sum_{\abs{\N}=N}X/\Tr\P_{N,0}$ is an
average of $X$ over all choices of $N$ distinct nuclear spins.

\begin{figure}
  \centering
  \includegraphics{time_table.pdf}
  \caption{\footnotesize Net population of the $N$-body orbital
    excited states $\set{\ket{\N,\pm}}$ after interrogation of an
    initial mixed state $\rho_{N,0}=\P_{N,0}/\Tr\P_{N,0}$ for a
    reduced time $\tau_N\equiv t\Omega_0/\sqrt{N}$ (i.e.~with real
    time $t$) by a laser with bare Rabi frequency $\Omega_0$ which is
    resonant with the $N$-body excitation energy $\Delta_{NX}$.  Here
    $\P_{NX}$, defined in \eqref{eq:P_NX}, is a projector onto the
    space of the $N$-body orbital states $\ket{\N X}$.}
  \label{fig:time}
\end{figure}

Figure \ref{fig:time} shows the excited-state population
$\bk{\P_{N,\pm}\p{t}}$ for several occupation numbers $N$.  With the
exception of $N=2$, the asymmetric-state populations generally have an
initial short period of growth before falling back to
$\bk{\P_{N,-}}\approx1/2$.  This behavior can be understood by the
fact that for fixed $N>2$, any pair of Rabi frequencies
$\Omega_{\N_1,-},\Omega_{\N_2,-}$ with
$\Omega_{\N_1,-}\ne\Omega_{\N_2,-}$ are mutually incommensurate, which
implies that at times $t$ with
$\min_{\abs{\N}=N}\set{2t\Omega_{\N,-}}\gtrsim1$ the averaging in
\eqref{eq:P_N_pm_t} effectively becomes a pseudo-random sampling
average of $\cos x$ over values of $x$, so
$\bk{\cos\p{2t\Omega_{\N,-}}}_{\abs{\N}=N}\approx0$.  When $N=2$, the
asymmetric-state Rabi frequencies essentially take on the same values
as the symmetric-state ones (see figure \ref{fig:coefficients}); the
behavior of asymmetric-state population dynamics for $N=2$ can
therefore be understood by the following discussion of symmetric-state
population dynamics.

To understand the periodic collapse and revival of symmetric-state
populations in figure \ref{fig:time}, we observe from \eqref{eq:O_N+}
that the symmetric-state phases in \eqref{eq:rho_N_pm_t} and
\eqref{eq:P_N_pm_t} take the form
\begin{align}
  2t\Omega_{\N,+} = \tau_N \sum_{\mu\in\N} 2\mu
  &&
  \t{with}
  &&
  \tau_N \equiv \f{t\Omega_0}{\sqrt{N}},
  \label{eq:2tO_N+}
\end{align}
where for fermionic atoms with half-integer nuclear spin, $2\mu$ is
always an odd integer, which implies that the sum in \eqref{eq:2tO_N+}
is an integer with the same parity (i.e.~even/odd) as the occupation
number $N$ (i.e.~the number of elements in $\N$).  At reduced times
$\tau_N=n\pi$ with integer $n$, therefore, if the occupation number
$N$ is even then all phases $2t\Omega_{\N,+}$ are integer multiples of
$2\pi$, which leads to a collapse of the excited-state populations as
$\left.\rho_N^{(\pm)}\p{t}\right|_{\tau_N=n\pi} = \rho_N^{(\pm)}\p{0}
= \rho_{N,0}$.  If the occupation $N$ is odd, meanwhile, then the
phases $2t\Omega_{\N,+}$ are all odd (even) integer multiples of $\pi$
for odd (even) $n$.  This alignment of phases implies a complete
population transfer to the excited state
$\rho_{N,-}\equiv\P_{N,-}/\Tr\P_{N,-}$ for odd $n$, and a collapse
back to the orbital ground state $\rho_{N,0}$ for even $n$, precisely
as observed in figure \ref{fig:time}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Summary and outlook}
\label{sec:summary}

Current 3-D optical lattice experiments with fermionic AEAs are
capable of operating in the low-temperature, high-density,
strongly-interacting limit where inter-atomic interactions set the
dominant energy scale governing system dynamics.  For AEAs with total
nuclear spin $I$ and $N=2I+1$ nuclear spin states, these interactions
exhibit an exotic SU($N$) symmetry which is of great interest for
near-term quantum simulations of SU($N$) spin models and lattice field
theories.  Working in the deep-lattice limit, we have derived a
low-energy effective theory of such atoms.  This theory exhibits
emergent multi-body interactions that inherit the SU($N$) symmetry of
the atoms' bare, two-body interactions.  The SU($N$) symmetry of
$M$-body Hamiltonians in the effective theory has allowed us to
express them in a simple form, and fully characterize their
corresponding eigenstates and spectra.  Capitalizing on the extreme
precision of state-of-the-art clock spectroscopy, we have tested
spectral predictions of our theory against direct experimental
measurements of the many-body ${}^{87}$Sr excitation spectrum.  This
comparison shows good agreement between theory and experiment, clearly
demonstrating the need to consider multi-body effects for
understanding the low-energy physics of high density AEA samples on a
3-D lattice.  Finally, we analyzed the many-body orbital-state
dynamics of multiply-occupied lattice sites prepared in a nuclear spin
mixture and interrogated via Rabi spectroscopy.  This analysis is
useful for future experimental probes of many-body state structures,
as well as for the preparation of long-lived states with multi-partite
entanglement (i.e.~$\ket{\N,\pm}$) which may be used as a resource to
perform quantum information processing tasks.

Despite the nominal success of our low-energy effective theory in
reproducing experimental observations, there remains room for
improvement in the form of controlled, systematic treatment of
higher-order and tunneling processes.
% These efforts will be necessary to go beyond the deep-lattice limit
% and study inter-site orbital-state and nuclear spin exchange
% dynamics.  Nonetheless, our work makes a major step towards the
% experimental investigation of multi-body SU($N$) physics.
Nonetheless, our work makes a major step towards the experimental
investigation of multi-body SU($N$) physics, providing the necessary
framework for future studies going beyond the deep-lattice limit to
realize multi-body super-exchange dynamics and orbital SU($N$) quantum
magnetism with AEAs.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Acknowledgements}

We acknowledge helpful discussions with R.~B.~Hutson, G.~E.~Marti,
S.~L.~Campbell, J.~Ye, P.~Julienne, J.~P.~D'Incao, C.~Kennedy, and
L.~Radzihovsky; and in particular close correspondence and technical
contributions from A.~Goban.  This work is supported by NIST, DARPA,
W911NF-16-1-0576 through ARO, AFOSR, AFOSR-MURI, NSF-1734006, NASA,
and CTQM.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Appendices}
\appendix

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Derivation of the effective Hamiltonian expansion}
\label{sec:eff_derivation}

Suppose we have a Hamiltonian $H_0$ on a Hilbert space
$\H=\G_0\oplus\E_0$ for a zero-energy manifold $\G_0$ decoupled from a
positive-energy manifold $\E_0$, and that we perturb $H_0$ by an
operator $V$ which weakly couples $\G_0$ and $\E_0$.  For all
$\ket{\psi_g}\in\G_0$ and $\ket{\psi_e},\ket{\chi_e}\in\E_0$, we have
$\bk{\psi_g|H_0|\psi_g}=0$, $\bk{\psi_e|H_0|\psi_g}=0$, and
$\abs{\bk{\psi_e|V|\psi_g}}\ll\bk{\chi_e|H_0|\chi_e}$.  The net
Hamiltonian $H=H_0+V$ will naturally admit a decomposition of the
Hilbert space as $\H=\G\oplus\E$ for a subspace $\G$ which is spanned
by the low-energy eigenstates of $H$ and has the same dimension as
$\G_0$, i.e.~$\abs{\G}=\abs{\G_0}$.

We can perform a canonical transformation between $\G$ and $\G_0$
which yields an {\it effective Hamiltonian} $H_{\t{eff}}$ on $\G_0$
that reproduces the spectrum of $H$ on $\G$
\cite{bravyi2011schrieffer}.  Given an eigenbasis
$\set{\ket{\alpha_0}}$ for $H_0$ on $\G_0$ and $\set{\ket\alpha}$ for
$H$ on $\G$, this transformation is implemented by a unitary $U$ for
which $\ket{\alpha_0}=U\ket\alpha$ and $U\to\1$ as $\norm{V}\to0$.
The effective Hamiltonian is then simply
\begin{align}
  H_{\t{eff}} = U H U^\dag.
  \label{eq:H_eff}
\end{align}
The prescription in \eqref{eq:H_eff} for constructing an effective
Hamiltonian is commonly known as a Schieffer-Wolff transformation
\cite{schrieffer1966relation}.  Unitaries $U$ which follow this
prescription are not unique, and different choices of $U$ amount to
different realizations of the Schieffer-Wolff transformation.  In
ref.~\cite{bravyi2011schrieffer}, the authors construct the unique
operator $S$ which generates a {\it direct} or {\it minimal} rotation
$U_{\t{min}}=e^S$ between $\G$ and $\G_0$, and use this construction
to expand \eqref{eq:H_eff} as a perturbative series in $V$.  The
rotation $U_{\t{min}}$ is minimal in the sense that it minimizes the
distance of candiate unitaries $U$ from the identity $\1$ with respect
to the Euclidian operator norm\footnote{The Euclidean operator norm is
  also known as the $L_{2,2}$, Hilbert-Schmidt, or Frobenius norm.}
$\norm{X}_{\t{E}}\equiv\sqrt{\Tr\p{X^\dag X}}$.  This rotation is
determined uniquely by enforcing
\begin{enumerate*}
\item that the generator $S$ is strictly block-off-diagonal with
  respect to $\G_0$ and $\E_0$,
\item that the norm $\norm{S}_{\t{E}}<\pi/2$, and
\item that the block-off-diagonal parts of \eqref{eq:H_eff} are zero.
\end{enumerate*}

To summarize the solution in ref.~\cite{bravyi2011schrieffer}, the
effective Hamiltonian $H_{\t{eff}}$ induced by a direct rotation can
be expanded as
\begin{align}
  H_{\t{eff}} = \sum_{p\ge0} H_{\t{eff}}^{(p)},
  \label{eq:H_eff_expansion}
\end{align}
where $H_{\t{eff}}^{(p)}$ is order $p$ in $V$.  Letting
$\Q_0\equiv1-\P_0$ denote the projector onto $\E_0$ and $X$ denote any
operator on $\H$, we define the superoperators
\begin{align}
  \D X \equiv \P_0 X \P_0 + \Q_0 X \Q_0,
  &&
  \O X \equiv \P_0 X \Q_0 + \Q_0 X \P_0,
\end{align}
which select out the diagonal ($\D$) and off-diagonal ($\O$) parts of
$X$ with respect to $\G_0$ and $\E_0$, and
\begin{align}
  \L X \equiv \sum_{\alpha,\beta}
  \f{\op\alpha \O X \op\beta}{E_\alpha - E_\beta}
  &&
  \t{where}
  &&
  H_0 = \sum_\alpha E_\alpha \op\alpha.
\end{align}
The first few terms of the expansion in \eqref{eq:H_eff_expansion} are
then, as derived in ref.~\cite{bravyi2011schrieffer},
\begin{align}
  H_{\t{eff}}^{(0)} = \P_0 H_0 \P_0,
  &&
  H_{\t{eff}}^{(1)} = \P_0 V \P_0,
  \label{eq:H_eff_0_1}
\end{align}
\begin{align}
  H_{\t{eff}}^{(2)} = - \f12 \P_0 \sp{\O V,\L V} \P_0,
  &&
  H_{\t{eff}}^{(3)} = \f12 \P_0 \sp{\O V, \L \sp{\D V, \L V}} \P_0.
\end{align}
Exploiting the fact that in our case $\bk{\psi|H_0|\psi}=0$ for all
$\ket\psi\in\G_0$, we let $\B_0\p{\E_0}$ denote an eigenbasis of $H_0$
for $\E_0$ and define the operator
\begin{align}
  \I \equiv \sum_{\ket\alpha\in\B_0\p{\E_0}}\f{\op\alpha}{E_\alpha},
\end{align}
which sums over projections onto excited states with corresponding
energetic suppression factors.  We then expand
\begin{align}
  \L X = \O\p{\L X} = \Q_0 \L X \P_0 + \P_0 \L X \Q_0
  = \I X \P_0 - \P_0 X \I,
\end{align}
which simplifies the expression for $H_{\t{eff}}^{(2)}$ as
\begin{align}
  H_{\t{eff}}^{(2)}
  = - \f12 \P_0 \p{\sp{\O V, \I V \P_0}
    - \sp{\O V, \P_0 V \I}} \P_0
  = - \P_0 V \I V \P_0.
  \label{eq:H_eff_2}
\end{align}
Working toward a similar expansion for $H_{\t{eff}}^{(3)}$, we compute
\begin{align}
  \sp{\D V, \L V}
  = \sp{\D V, \I V \P_0} - \sp{\D V, \P_0 V \I}
  = \O\p{V \I V} - \I V \P_0 V \P_0 - \P_0 V \P_0 V \I,
\end{align}
and in turn
\begin{align}
  H_{\t{eff}}^{(3)}
  &= \f12 \P_0 \p{\sp{\O V, \I \sp{\D V, \L V} \P_0}
    - \sp{\O V, \P_0 \sp{\D V, \L V} \I}} \P_0 \\
  &= \f12 \P_0 \p{V \I \sp{\D V, \L V}
    + \sp{\D V, \L V} \I V} \P_0 \\
  &= \P_0 V \I V \I V \P_0
  - \f12 \P_0 V \I^2 V \P_0 V \P_0
  - \f12 \P_0 V \P_0 V \I^2 V \P_0 \\
  &= \P_0 V \I V \I V \P_0
  - \f12 \sp{\P_0 V \P_0, \P_0 V \I^2 V \P_0}_+,
  \label{eq:H_eff_3}
\end{align}
where $\sp{X,Y}_+\equiv XY+YX$.  The expressions in
\eqref{eq:H_eff_0_1}, \eqref{eq:H_eff_2}, and \eqref{eq:H_eff_3}
complete the derivation for our expansion of the effective interaction
Hamiltonian $H_{\t{int}}^{\t{eff}}$ in \eqref{eq:H_int_eff} through
third order.  In the case of ultracold atoms on a lattice, the
motional ground-state subspace $\G_0$ actually contains many internal
atomic states with different energies.  Nonetheless, the total Hilbert
space is completely separable into uncoupled subspaces associated with
each symmetrized many-body internal atomic state.  One can therefore
diagonalize the interaction Hamiltonian with respect to these internal
states and derive an effective theory within each of the corresponding
subspaces, in each case setting the appropriate ground-state energy to
zero.  This procedure is equivalent to simultaneously calculating the
effective Hamiltonian $H_{\t{int}}^{\t{eff}}$ for all internal states
via the prescriptions we have provided, but letting $E_\alpha$ denote
only the motional excitation energy of states
$\ket{\alpha_0}\in\B_0\p{\E_0}$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Diagram counting and symmetry factors}
\label{sec:diagrams}

The fact that we include factors of $1/2$ from the bare two-body
interaction Hamiltonian $H_{\t{int}}$ in the definition of diagrams
implies that $p$-vertex diagrams acquire a factor of $1/2^p$.  In
practice, however, these factors are exactly cancelled out by
corresponding symmetry factors in all $M$-body diagrams with $M>2$.
As an illustrative example, consider the second-order effective
Hamiltonian $H_{\t{int}}^{(2)}$ in \eqref{eq:H_int_1_2}, which
expanded in full reads
\begin{align}
  H_{\t{int}}^{(2)} = -\sum_{m+n>0} \P_0
  \p{\f12 K_{mn} G^{st}_{s't'}
    \c_{\rho s'}^\dag \c_{\sigma t'}^\dag \c_{n\sigma t} \c_{m\rho s}}
  \f1{E_{mn}}
  \p{\f12 K_{mn} G^{qr}_{q'r'}
    \c_{m\mu q'}^\dag \c_{n\nu r'}^\dag \c_{\nu r} \c_{\mu q}} \P_0.
  \label{eq:H_int_2_full}
\end{align}
The three-body terms in this Hamiltonian have
$\abs{\set{\mu,\nu,\rho,\sigma}}=3$ and only one virtually excited
atom.  The non-vanishing three-body terms must therefore either have
$\rho\in\set{\mu,\nu}$ and contain a factor of the form
$\c_X^\dag \c_{\sigma t'}^\dag \c_{\sigma t} \c_Y$, or have
$\sigma\in\set{\mu,\nu}$ with a factor of the form
$\c_{\rho s'}^\dag \c_X^\dag \c_Y \c_{\rho s}$, where the labels $X,Y$
both address whichever nuclear spin (i.e.~$\mu$ or $\nu$) was excited
in the corresponding term.  Diagrammatically, we have terms of the
form
\begin{align}
  \begin{tikzpicture}
    \begin{feynman}
      \vertex (v1);
      \vertex[above left = of v1] (f1) {$\mu q$};
      \vertex[below left = of v1] (f2) {$\nu r$};
      \vertex[right = of v1] (vm);
      \vertex[right = of vm] (v2);
      \vertex[above = of vm] (f3) {$Z$};
      \vertex[below = of vm] (f4) {$\sigma t$};
      \vertex[below right = of v2] (f5) {$\sigma t'$};
      \vertex[above right = of v2] (f6) {$X$};
      \diagram* {
        (f1) -- (v1) -- (f3),
        (f2) -- (v1) --  [scalar, edge label=$Y$] (v2),
        (f4) -- (v2) -- (f5),
        (v2) -- (f6), };
    \end{feynman}
  \end{tikzpicture}
  &&
  \t{and}
  &&
  \begin{tikzpicture}
    \begin{feynman}
      \vertex (v1);
      \vertex[above left = of v1] (f1) {$\mu q$};
      \vertex[below left = of v1] (f2) {$\nu r$};
      \vertex[right = of v1] (vm1);
      \vertex[right = of vm1] (vm2);
      \vertex[right = of vm2] (v2);
      \vertex[above = of vm1] (f3) {$Z$};
      \vertex[above = of vm2] (f4) {$\rho s$};
      \vertex[above right = of v2] (f5) {$\rho s'$};
      \vertex[below right = of v2] (f6) {$X$};
      \diagram* {
        (f1) -- (v1) -- (f3),
        (f2) -- (v1) --  [scalar, edge label=$Y$] (v2),
        (f4) -- (v2) -- (f5),
        (v2) -- (f6), };
    \end{feynman}
  \end{tikzpicture}.
  \label{eq:rho_sigma}
\end{align}
Observing that
$\c_{\rho s'}^\dag \c_X^\dag \c_Y \c_{\rho s} = \c_X^\dag \c_{\rho
  s'}^\dag \c_{\rho s} \c_Y$, however, it is clear that both of the
terms represented in \eqref{eq:rho_sigma} are equal up to the
re-indexing $\p{\sigma,t,t'}\leftrightarrow\p{\rho,s,s'}$.  There is
therefore a symmetry factor of $2$ associated with the second vertex
of the diagrams in \eqref{eq:rho_sigma}, which cancels out with the
explicit factor of $1/2$ at that vertex, i.e.~the first factor of
$1/2$ in \eqref{eq:H_int_2_full}.  A symmetry factor of essentially
identical origin appears at every vertex with an ``incoming'' virtual
state, as in e.g.~the second and third vertices of
\begin{align}
  \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[below right = of v1] (v2);
        \vertex[above right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[left = of v1] (f2);
        \vertex[below left = of v2] (f3);
        \vertex[above right = of v3] (f4);
        \vertex[right = of v3] (f5);
        \vertex[below right = of v2] (f6);
        \diagram* {
          (f1) -- (v1) --[scalar] (v3) -- (f4),
          (f2) -- (v1) --[scalar] (v2) --[scalar] (v3) -- (f5),
          (f3) -- (v2) -- (f6), };
      \end{feynman}
    \end{tikzpicture}}
  &&
  \t{and}
  &&
  \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[below right = of v1] (v2);
        \vertex[below right = of v2] (v3);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[below left = of v2] (f3);
        \vertex[below left = of v3] (f4);
        \vertex[above right = of v1] (f5);
        \vertex[above right = of v2] (f6);
        \vertex[above right = of v3] (f7);
        \vertex[below right = of v3] (f8);
        \diagram* {
          (f1) -- (v1) -- (f5),
          (f2) -- (v1) -- (f5),
          (v1) --[scalar] (v2) -- (v3) -- (f8),
          (f3) -- (v2) -- (f6),
          (f4) -- (v3) -- (f7),
        };
      \end{feynman}
    \end{tikzpicture}},
  \label{eq:diagram_example_1}
\end{align}
or the last vertex of
\begin{align}
  \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[below right = 2.5em of v1] (v3);
        \vertex[below left = 1.7em of v3] (v2);
        \vertex[above left = of v1] (f1);
        \vertex[left = of v1] (f2);
        \vertex[left = of v2] (f3);
        \vertex[below left = of v2] (f4);
        \vertex[above right = of v1] (f5);
        \vertex[above right = of v3] (f6);
        \vertex[below right = of v3] (f7);
        \vertex[below right = of v2] (f8);
        \diagram* {
          (f1) -- (v1) -- (f5),
          (f2) -- (v1) --[scalar] (v3) -- (f6),
          (f3) -- (v2) -- (v3) -- (f7),
          (f4) -- (v2) -- (f8), };
      \end{feynman}
    \end{tikzpicture}}.
  \label{eq:diagram_example_2}
\end{align}
We can thus account for cancellations of $1/2$ at all vertices except
those which address two ``initial'' ground-state atoms, as in the
first vertex of the diagrams in \eqref{eq:rho_sigma} and
\eqref{eq:diagram_example_1}, or the first two vertices of the diagram
in \eqref{eq:diagram_example_2}.  For such vertices, there are two
possibilities: either
\begin{enumerate}
\item both edges leaving the vertex in question (i.e.~leaving to the
  right) terminate at different vertices, as in the examples above, or
  \label{enum:different_vertices}
\item both edges leaving the vertex in question terminate at the same
  vertex, as in for example the first vertex of
  \label{enum:same_vertex}
  \begin{align}
    \shrink{
      \begin{tikzpicture}
        \begin{feynman}
          \vertex (v1);
          \vertex[right = of v1] (v2);
          \vertex[below right = of v2] (v3);
          \vertex[above left = of v1] (f1);
          \vertex[below left = of v1] (f2);
          \vertex[below left = of v3] (f3);
          \vertex[above right = of v2] (f4);
          \vertex[above right = of v3] (f5);
          \vertex[below right = of v3] (f6);
          \diagram* {
            (f1) -- (v1) --[half left, scalar] (v2) -- (f4),
            (f2) -- (v1) --[half right, scalar] (v2)
            --[scalar] (v3) -- (f5),
            (f3) -- (v3) -- (f6), };
        \end{feynman}
      \end{tikzpicture}}.
    \label{eq:diagram_example_3}
  \end{align}
\end{enumerate}
In the former case, \ref{enum:different_vertices}, the vertex in
question has an associated symmetry factor of 2 to account for the
possibility of a nuclear spin exchange at that vertex.  Considering
again our example of the second-order effective Hamiltonian
$H_{\t{int}}^{(2)}$ in \eqref{eq:H_int_2_full}, the non-vanishing
three-body terms must either have $m=0$ and contain the factor
$\c_{\mu q'}^\dag \c_{n\nu r'}^\dag \c_{\nu r} \c_{\mu q}$, or have
$n=0$ with the factor
$\c_{m\mu q'}^\dag \c_{\nu r'}^\dag \c_{\nu r} \c_{\mu q}$, which
diagrammatically translates to
\begin{align}
  \begin{tikzpicture}
    \begin{feynman}
      \vertex (v1);
      \vertex[above left = of v1] (f1) {$\mu q$};
      \vertex[below left = of v1] (f2) {$\nu r$};
      \vertex[right = of v1] (vm);
      \vertex[right = of vm] (v2);
      \vertex[above = of vm] (f3) {$\mu q'$};
      \vertex[below = of vm] (f4) {$\sigma t$};
      \vertex[below right = of v2] (f5) {$\sigma t'$};
      \vertex[above right = of v2] (f6) {$\nu r''$};
      \diagram* {
        (f1) -- (v1) -- (f3),
        (f2) -- (v1) --  [scalar, edge label=$n\nu r'$] (v2),
        (f4) -- (v2) -- (f5),
        (v2) -- (f6), };
    \end{feynman}
  \end{tikzpicture}
  &&
  \t{or}
  &&
  \begin{tikzpicture}
    \begin{feynman}
      \vertex (v1);
      \vertex[above left = of v1] (f1) {$\mu q$};
      \vertex[below left = of v1] (f2) {$\nu r$};
      \vertex[right = of v1] (vm);
      \vertex[right = of vm] (v2);
      \vertex[above = of vm] (f3) {$\nu r'$};
      \vertex[below = of vm] (f4) {$\sigma t$};
      \vertex[below right = of v2] (f5) {$\sigma t'$};
      \vertex[above right = of v2] (f6) {$\mu q''$};
      \diagram* {
        (f1) -- (v1) -- (f3),
        (f2) -- (v1) --  [scalar, edge label=$m\mu q'$] (v2),
        (f4) -- (v2) -- (f5),
        (v2) -- (f6), };
    \end{feynman}
  \end{tikzpicture}.
  \label{eq:mu_nu}
\end{align}
These terms are equal up to the re-indexing
$\p{\nu,r,r',r'',n}\leftrightarrow\p{\mu,q,q',q'',m}$, which implies
that there is a symmetry factor of $2$ associated with the first
vertex of the diagrams in \eqref{eq:mu_nu}.  A symmetry factor of
identical origin is associated with the first vertex of the diagrams
in \eqref{eq:diagram_example_1}, and the first two vertices of the
diagram in \eqref{eq:diagram_example_2}.

The final case we must consider is \ref{enum:same_vertex}, which
occurs in the first vertex of \eqref{eq:diagram_example_3}.  In this
case, the symmetry factor of $2$ which appears in case
\ref{enum:different_vertices} to account for the possibility of a
nuclear spin exchange simply gets ``pushed forward'' to the vertex at
which the two nuclear spins in question part ways, e.g.~to account for
the two possibilities
\begin{align}
  \begin{tikzpicture}
    \begin{feynman}
      \vertex (v1);
      \vertex[right = of v1] (v2);
      \vertex[below right = of v2] (v3);
      \vertex[above left = of v1] (f1) {$\mu r$};
      \vertex[below left = of v1] (f2) {$\nu s$};
      \vertex[below left = of v3] (f3) {$\rho t$};
      \vertex[above right = of v2] (f4) {$\mu r''$};
      \vertex[above right = of v3] (f5) {$\nu s'''$};
      \vertex[below right = of v3] (f6) {$\rho t'$};
      \diagram* {
        (f1) -- (v1)
        --[half left, scalar, edge label=$\ell\mu r'$] (v2)
        -- (f4),
        (f2) -- (v1)
        --[half right, scalar, edge label'=$m\nu s'$] (v2)
        --[scalar, edge label=$n\nu s''$] (v3)
        -- (f5),
        (f3) -- (v3) -- (f6), };
    \end{feynman}
  \end{tikzpicture}
  &&
  \t{and}
  &&
  \begin{tikzpicture}
    \begin{feynman}
      \vertex (v1);
      \vertex[right = of v1] (v2);
      \vertex[below right = of v2] (v3);
      \vertex[above left = of v1] (f1) {$\mu r$};
      \vertex[below left = of v1] (f2) {$\nu s$};
      \vertex[below left = of v3] (f3) {$\rho t$};
      \vertex[above right = of v2] (f4) {$\nu s''$};
      \vertex[above right = of v3] (f5) {$\mu r'''$};
      \vertex[below right = of v3] (f6) {$\rho t'$};
      \diagram* {
        (f1) -- (v1)
        --[half left, scalar, edge label=$\ell\mu r'$] (v2)
        -- (f4),
        (f2) -- (v1)
        --[half right, scalar, edge label'=$m\nu s'$] (v2)
        --[scalar, edge label=$n\mu r''$] (v3)
        -- (f5),
        (f3) -- (v3) -- (f6), };
    \end{feynman}
  \end{tikzpicture},
  \label{eq:mu_nu_loop}
\end{align}
which are equal up to the re-indexing
$\p{\nu,s,s',s'',s''',m}\leftrightarrow\p{\mu,r,r',r'',r''',\ell}$.
The arguments for a symmetry factor in cases
\ref{enum:different_vertices} and \ref{enum:same_vertex} fail only if
the two nuclear spins in question take identical paths through the
internal vertices of a diagram, such that there is no meaningful sense
in which two diagrams can be said to differ by a nuclear spin
exchange, as in \eqref{eq:mu_nu} and \eqref{eq:mu_nu_loop}.  If two
atoms take identical paths through the internal vertices of a diagram,
however, then they have only participated in a two-body process, as in
\begin{align}
  \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v);
        \vertex[above left = of v] (f1);
        \vertex[below left = of v] (f2);
        \vertex[above right = of v] (f3);
        \vertex[below right = of v] (f4);
        \diagram* {
          (f1) -- (v),
          (f2) -- (v),
          (v) -- (f3),
          (v) -- (f4) };
      \end{feynman}
    \end{tikzpicture}},
  &&
  \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[right = of v1] (v2);
        \vertex[above right = of v2] (f3);
        \vertex[below right = of v2] (f4);
        \diagram* {
          (f1) -- (v1),
          (f2) -- (v1),
          (v2) -- (f3),
          (v2) -- (f4),
          (v1) --[scalar, half left] (v2),
          (v1) --[scalar, half right] (v2) };
      \end{feynman}
    \end{tikzpicture}},
  &&
  \shrink{
    \begin{tikzpicture}
      \begin{feynman}
        \vertex (v1);
        \vertex[above left = of v1] (f1);
        \vertex[below left = of v1] (f2);
        \vertex[right = of v1] (v2);
        \vertex[right = of v2] (v3);
        \vertex[above right = of v3] (f3);
        \vertex[below right = of v3] (f4);
        \diagram* {
          (f1) -- (v1),
          (f2) -- (v1),
          (v3) -- (f3),
          (v3) -- (f4),
          (v1)
          --[scalar, half left] (v2)
          --[scalar, half left] (v3),
          (v1)
          --[scalar, half right] (v2)
          --[scalar, half right] (v3) };
      \end{feynman}
    \end{tikzpicture}},
  &&
  \cdots.
\end{align}
After summing over all free indices, therefore, all two-body diagrams
have a remaining factor of $1/2$ from the first vertex.  In all
connected $M$-body diagrams with $M>2$, meanwhile, every factor of
$1/2$ can be identified one-to-one with a corresponding symmetry
factor of $2$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Effective coupling constants in a lattice}
\label{sec:renormalization}

Due to our choice of renormalization scheme in section
\ref{sec:two_body}, the interaction energies prescribed by our
low-energy effective theory for multiply-occupied lattice sites are
not given directly by the coupling constants $G_X$ defined by the
free-space scattering lengths $a_X$ in \eqref{eq:couplings}.  Instead,
we must first compute effective coupling constants
$G_X^{\t{lattice}}\p{\U}$ in a lattice with depth $\U$, and in turn
use the effective coupling constants to compute interaction energies.
As the renormalization procedure $G_X\to G_X^{\t{lattice}}$ is
identical for all coupling constants, we henceforth drop the subscript
$X\in\set{\g\g,\e\g^-,\e\g^+,\e\e}$ on coupling constants $G_X$ in the
remainder of this appendix.  To further simplify notation, we will
also neglect the explicit dependence of parameters on the lattice
depth $\U$, which we generally keep fixed.

Proper calculations of the interaction energy of two ultracold
fermions in an optical lattice were performed in
refs.~\cite{buchler2010microscopic} and \cite{wall2013strongly} using
a two-channel model of a Feshbach resonance, yielding prescriptions
for computing effective coupling constants in a lattice from
free-space interaction parameters.  These calculations, however, are
both analytically and numerically involved.  We therefore instead opt
to use a modified version of the considerably simpler single-channel
calculation in ref.~\cite{busch1998two} of the interaction energy of
two ultracold atoms in a harmonic trap.  Our approach is equivalent to
the calculation of Hubbard parameters performed in
ref.~\cite{dickerscheid2005feshbach}, and has been demonstrated to
reproduce correct results in the limit of a deep lattice (compared to
the lattice photon recoil energy) and small positive scattering
lengths (compared to the effective harmonic oscillator length)
\cite{buchler2010microscopic}.

The exact result in Eq. 16 of ref.~\cite{busch1998two} for the
interaction energy of two ultracold atoms in a harmonic oscillator
with angular trap frequency $\omega$ can be written in the form
\begin{align}
  \p{G_{\t{free}} K_{\t{HO}}/\omega}^{-1}
  = \f{\sqrt{\pi}~ \Gamma\p{-G_{\t{HO}}K_{\t{HO}}/2\omega}}
  {\Gamma\p{-G_{\t{HO}}K_{\t{HO}}/2\omega-1/2}},
  &&
  K_{\t{HO}} \equiv \int d^3x~ \abs{\phi_0^{\t{HO}}}^4,
  \label{eq:harmonic_gamma}
\end{align}
where $G_{\t{HO}}$ is an effective coupling constant in the harmonic
trap, $\phi_0^{\t{HO}}\p{x}$ is the corresponding single-particle
ground-state wavefunction, and $\Gamma$ is the gamma function.  The
expression in \eqref{eq:harmonic_gamma} can be solved numerically as
is, or expanded about $G_{\t{HO}}K_{\t{HO}}/\omega=0$ to get
\begin{align}
  G_{\t{free}}^{-1} = G_{\t{HO}}^{-1}
  \sum_{n=0}^\infty c_n \p{G_{\t{HO}}K_{\t{HO}}/\omega}^n,
  \label{eq:harmonic_series}
\end{align}
where the first few coefficients are
\begin{align}
  c_0 = 1,
  &&
  c_1 = 1 - \ln 2,
  &&
  c_2 = -\f{\pi^2}{24} - \ln 2 + \f12 \p{\ln 2}^2.
\end{align}
The series in \eqref{eq:harmonic_series} can in turn be inverted to
solve for $G_{\t{HO}}$ with an expansion of the form
\begin{align}
  G_{\t{HO}} = G_{\t{free}}
  \sum_{n=0}^\infty \tilde c_n \p{G_{\t{free}}K_{\t{HO}}/\omega}^n,
  \label{eq:inverted_series}
\end{align}
where if we truncate the series in \eqref{eq:harmonic_series} at
$n=2$, the first few coefficients of \eqref{eq:inverted_series} are
\begin{align}
  \tilde c_0 = 1,
  &&
  \tilde c_1 = 1 - \ln 2,
  &&
  \tilde c_2 = -\f{\pi^2}{24} - \ln 2 + \f12 \p{\ln 2}^2
  + \p{1 - \ln 2}^2.
  \label{eq:inverted_coefficients}
\end{align}
The coefficients $\tilde c_n$ thus found are consistent with the
coefficients $c_2^{(n+1)}$ reported in table 1 of
ref.~\cite{johnson2012effective}, in which the authors compute the
first few terms of the two-body Hamiltonian $H_2$ directly as
expressed in \eqref{eq:H_2_expansion} by using a renormalization
scheme which subtracts off divergences term by term.

All of the above results are exact for two atoms in a harmonic
oscillator interacting via $s$-wave scattering.  In order to adapt
these results for a lattice, we expand the lattice potential about a
lattice site centered at $x=(0,0,0)$ as
\begin{align}
  \U\sin^2\p{k_L\cdot x} \approx \U \p{k_L^x k_L^y k_L^z}^{2/3} x^2
  = \f12 m_A \omega_{\t{eff}}^2 x^2,
  &&
  \omega_{\t{eff}} \equiv \sqrt{2~ \U k_L^2 / m_A},
\end{align}
where $k_L=\p{k_L^x,k_L^y,k_L^z}$ is the lattice wavenumber, $m_A$ is
the atomic mass, and $\omega_{\t{eff}}$ is an effective angular
harmonic trap frequency.  We then use $\omega_{\t{eff}}$ in place of
$\omega$ in \eqref{eq:harmonic_gamma}, and use an overlap integral $K$
computed with the ground-state wavefunctions $\phi_0$ in a lattice
rather than those in a harmonic oscillator.  We retrieve free-space
$s$-wave scattering lengths $a_{\t{free}}$ for ${}^{87}$Sr from
ref.~\cite{zhang2014spectroscopic} to determine the free-space
coupling constants $G_{\t{free}} \equiv \p{4\pi/m_A} a_{\t{free}}$.
This procedure yields an effective coupling constant $G_{\t{lattice}}$
given by
\begin{align}
  \p{G_{\t{free}} K/\omega_{\t{eff}}}^{-1}
  = \f{\sqrt{\pi}~ \Gamma\p{-G_{\t{lattice}}K/2\omega_{\t{eff}}}}
  {\Gamma\p{-G_{\t{lattice}}K/2\omega_{\t{eff}}-1/2}},
\end{align}
with a solution
\begin{align}
  G_{\t{lattice}} = G_{\t{free}}
  \sum_{n=0}^\infty \tilde c_n \p{G_{\t{free}}K/\omega_{\t{eff}}}^n,
\end{align}
where the first few coefficients are provided in
\eqref{eq:inverted_coefficients}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Momentum-dependent $s$-wave interactions}
\label{sec:momentum_dependence}

In addition to the renormalization of coupling constants discussed in
Appendix \ref{sec:renormalization}, computing two-body interaction
energies $E_{NX}^{(2)}$ at third order in the low-energy effective
theory requires accounting for the contribution of momentum-dependent
$s$-wave interactions.  At next-to-leading order in the relative
momentum $k$ between two atoms, the effective momentum-dependent
scattering length $a_{\t{eff}}$ is given in terms of the zero-momentum
scattering length $a$ by \cite{blume2002fermi, chin2010feshbach}
\begin{align}
  \f1{a_{\t{eff}}}
  = \f1a - \f12 r_{\t{eff}} k^2
  = \f1a\p{1 - \f12 r_{\t{eff}} a k^2},
\end{align}
which for $r_{\t{eff}} a k^2 \ll 1$, implies that
\begin{align}
  a_{\t{eff}}
  \approx a\p{1 + \f12 r_{\t{eff}} a k^2}
  = a + \f12 r_{\t{eff}} a^2 k^2.
  \label{eq:a_eff}
\end{align}
Here $r_{\t{eff}}$ is an effective range of $\O\p{k^2}$ interactions,
determined in atomic units by the scattering length $a$ and van der
Waals $C_6$ coefficient by \cite{flambaum1999analytical}
\begin{align}
  r_{\t{eff}} = \f13 \xi^{-2} \chi \p{1 - 2\chi + 2\chi^2} a,
  &&
  \t{where}
  &&
  \xi \equiv \f{\Gamma\p{3/4}}{\Gamma\p{1/4}},
  &&
  \chi \equiv \sqrt{2}~ \xi~ \f{\p{m_AC_6}^{1/4}}{a},
  \label{eq:effective_range}
\end{align}
and $\Gamma$ is the gamma function.  As $\chi\sim1$ for ${}^{87}$Sr,
the momentum-dependent correction to the effective scattering length
$a_{\t{eff}}$ is $\O\p{a^3}$ without an additional separation of
scales (i.e.~which could have occurred if we had $\chi\ll1$ or
$\chi\gg1$).  The momentum-independent $\O\p{k^0}$ contribution to
$a_{\t{eff}}$ in \eqref{eq:a_eff} gives rise to the bare two-body
interactions in \eqref{eq:H_int} by use of an unregularized contact
(i.e.~$\delta$-function) potential, while the momentum-dependent
$\O\p{k^2}$ term gives rise to the interaction Hamiltonian
\cite{blume2002fermi, johnson2012effective}
\begin{align}
  H_{\t{int}}'
  \equiv \f12 \sum {G'}^{qr}_{st} \int d^3 x~d^3y~\delta\p{z}
  \sp{\hat\psi_{\mu q}^\dag\p{x} \hat\psi_{\nu r}^\dag\p{y}}
  \hat k_z^2\sp{\hat\psi_{\nu t}\p{y} \hat\psi_{\mu s}\p{x}},
  \label{eq:H_int_primed}
\end{align}
where
\begin{align}
  z \equiv x - y,
  &&
  \hat k_z^2
  \equiv -\f12 \p{\vec\nabla_z^2 + \lvec\nabla_z^2},
\end{align}
and the primed couplings ${G'}^{qr}_{st}$ are defined similarly to
unprimed couplings $G^{qr}_{st}$ in \eqref{eq:couplings} and
\eqref{eq:coupling_tensor}, but with scattering lengths
$a\to r_{\t{eff}}a^2/2$ and the effective range defined by
\eqref{eq:effective_range} for each scattering length with the
appropriate $C_6$ coefficient.  We retrieve $C_6$ coefficients for
${}^{87}$Sr from the supplementary material of
ref.~\cite{zhang2014spectroscopic}.  The squared relative momentum
operator $k_z^2$ is represented by symmetrized left- and right-acting
derivative operators in order to preserve manifest Hermiticity of
$H_{\t{int}}'$.  At third order in the low-energy effective theory
developed in section \ref{sec:low_energy}, the bare momentum-dependent
interactions in \eqref{eq:H_2} yield only the effective two-body
Hamiltonian
\begin{align}
  H_2' \equiv \f12 K' \sum {G'}^{qr}_{st}
  \leq\c_{\mu s}^\dag \c_{\nu t}^\dag \c_{\nu r} \c_{\mu q},
  \label{eq:H_2_prime}
\end{align}
where, letting $\Re\sp{X}$ denote the real part of $X$,
\begin{align}
  K' \equiv \f12 \int d^3x~
  \Re\sp{\p{\phi_0^*}^2
    \p{\vec\nabla\phi_0\cdot\vec\nabla\phi_0
      - \phi_0\vec\nabla^2\phi_0}}.
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Bounds on theoretical uncertainties from inter-site effects}
\label{sec:error}

In our overview of the relevant one- and two-particle physics of
ultracold atoms on a lattice (section \ref{sec:overview}), we made two
approximations which introduce error into the low-energy effective
theory.  Both approximations concern the on-site locality of the
single- and two-body Hamiltonians: we assumed that
\begin{enumerate*}[label=\roman*)]
\item tunneling between lattice sites and
\item inter-site interactions are negligible.
\end{enumerate*}
These approximations are justified for single-particle motional ground
states of atoms in a deep lattice, but generally break down when
considering virtual states occupying highly excited motional levels,
whose spatial wavefunctions can span multiple lattice sites.
Nonetheless, we can place upper bounds on the magnitude of inter-site
corrections to the effective on-site interaction Hamiltonians by
treating tunneling and inter-site interactions of virtual excited
states perturbatively and assuming no energetic penalty for
nearest-neighbor hopping.  These bounds can be used to diagnose the
breakdown of the on-site effective theory, and signal when a more
careful consideration of inter-site effects is necessary to make
precise predictions about many-body spectra and dynamics.

If we still assume negligible overlap between single-particle
ground-state wavefunctions in different lattice sites but consider
nearest-neighbor wavefunction overlaps of states with motional
excitations, our one-body and bare two-body Hamiltonians become
\begin{align}
  H_0 = \sum E_n \c_{in\mu s}^\dag \c_{in\mu s}
  - \sum_{\substack{\bk{i,j}\\m,n>0}}
  \p{t_{mn} \c_{jn\mu s}^\dag \c_{im\mu s} + \t{h.c.}},
  \label{eq:H_0_neighbor}
\end{align}
and
\begin{align}
  H_{\t{int}} &= \f12 \sum K^{k\ell}_{mn} G^{qr}_{st}
  \c_{im\mu s}^\dag \c_{in\nu t}^\dag \c_{i\ell\nu r} \c_{ik\mu q}
  + \f12 \sum_{\substack{\bk{i,j}\\n>0}} G^{\mu q;\nu r}_{\rho s;\sigma t}
  \p{\K_n \c_{j,0,\rho s}^\dag \c_{j,0,\sigma t}^\dag
    \c_{j,0,\nu r} \c_{in\mu q} + \t{h.c.}} \nonumber \\
  &\quad + \f12 \sum_{\substack{\bk{i,j}\\m,n>0}}
  G^{\mu q;\nu r}_{\rho s;\sigma t}
  \p{\K_{mn} \c_{j,0,\rho s}^\dag \c_{j,0,\sigma t}^\dag
    \c_{in\nu r} \c_{im\mu q} +
    \tilde\K_{mn} \c_{in\rho s}^\dag \c_{j,0,\sigma t}^\dag
    \c_{j,0,\nu r} \c_{im\mu q} + \t{h.c.}},
  \label{eq:H_int_neighbor}
\end{align}
where $t_{mn}$ is a tunneling rate; $\K_n,\K_{mn},\tilde\K_{mn}$ are
inter-site spatial overlap integrals; $\t{h.c.}$ denotes a Hermitian
conjugate, i.e.~$\p{X+\t{h.c.}}\equiv \p{X+X^\dag}$; and $\bk{i,j}$
denotes the set of all lattice sites $i$ together with their adjacent
sites $j$.  Note that we have neglected terms in
\eqref{eq:H_int_neighbor} which involve more than two field operators
addressing states with motional excitations, as these terms will not
appear in the leading-order corrections to the effective on-site
interaction Hamiltonians.  We also still neglect terms which involve
products of atomic wavefunctions for motional ground states in
different lattice sites.

Diagrammatically representing matrix elements of $H_0$ and
$H_{\t{int}}$ which are off-diagonal in lattice site by a dot
(i.e.~$\bullet$) and marking lines which represent field operators
addressing neighboring lattice sites by a cross (i.e.~$+$ or $\times$,
depending on the line orientation), the dominant terms in the
effective theory which we previously neglected by assuming on-site
locality are
\begin{align}
  \begin{tikzpicture}
    \begin{feynman}
      \vertex (v1);
      \vertex[above left = of v1] (f1);
      \vertex[below left = of v1] (f2);
      \vertex[right = of v1] (v2);
      \vertex[above right = of v2] (f3);
      \vertex[below right = of v2] (f4);
      \diagram* {
        (f1) -- (v1),
        (f2) -- (v1),
        (v1) --[scalar, half left] (v2),
        (v1) --[scalar, half right] (v2),
        (v2) --[insertion = 0.5] (f3),
        (v2) --[insertion = 0.5] (f4) };
      \draw[fill=black] (v2) circle(0.8mm);
    \end{feynman}
  \end{tikzpicture}
  &\sim \gamma_2^{(2)} G^2,
  & \gamma_2^{(2)} &\equiv \sum_{n+m>0} \f{K_{mn}\K_{mn}}{E_{mn}},
  \label{eq:g_2_2} \\[1em]
  \begin{tikzpicture}
    \begin{feynman}
      \vertex (v1);
      \vertex[above left = of v1] (f1);
      \vertex[below left = of v1] (f2);
      \vertex[right = 4em of v1] (v2);
      \vertex[above right = of v1] (f3);
      \vertex[below left = of v2] (f4);
      \vertex[below right = of v2] (f5);
      \vertex[above right = of v2] (f6);
      \diagram* {
        (f1) -- (v1) -- (f3),
        (f2) -- (v1) --[scalar] (v2),
        (f4) --[insertion = 0.5] (v2) --[insertion = 0.5] (f5),
        (v2) --[insertion = 0.5] (f6), };
      \draw[fill=black] (v2) circle(0.8mm);
    \end{feynman}
  \end{tikzpicture}
  &\sim \gamma_{3,1}^{(2)} G^2,
  & \gamma_{3,1}^{(2)} &\equiv \sum_{n>0} \f{K_n\K_n}{E_n},
  \label{eq:g_3_1_2} \\[1em]
  \begin{tikzpicture}
    \begin{feynman}
      \vertex (v1);
      \vertex[above left = of v1] (f1);
      \vertex[below left = of v1] (f2);
      \vertex[right = 2em of v1] (v2);
      \vertex[right = 2em of v2] (v3);
      \vertex[above right = of v1] (f3);
      \vertex[below left = of v3] (f4);
      \vertex[below right = of v3] (f5);
      \vertex[above right = of v3] (f6);
      \diagram* {
        (f1) -- (v1) -- (f3),
        (f2) -- (v1) --[scalar] (v2) --[scalar, insertion = 0.5] (v3),
        (f4) --[insertion = 0.5] (v3) --[insertion = 0.5] (f5),
        (v3) --[insertion = 0.5] (f6), };
      \draw[fill=black] (v2) circle(0.8mm);
    \end{feynman}
  \end{tikzpicture}
  &\sim \gamma_{3,2}^{(2)} G^2,
  & \gamma_{3,2}^{(2)} &\equiv \sum_{n,m>0} \f{K_mt_{mn}K_n}{E_mE_n},
  \label{eq:g_3_2_2}
\end{align}
where we have identified, up to an assignment of coupling constants
$G$, the magnitude of all nonzero matrix elements of the diagrams with
respect to an eigenbasis of the on-site single-particle Hamiltonian
$H_0$ in \eqref{eq:H_0}.

The terms in \eqref{eq:g_2_2}-\eqref{eq:g_3_2_2} can be used to
estimate an upper bound on the magnitude of dominant corrections to
the spectrum of the low-energy theory from off-diagonal (i.e.~in
lattice site) matrix elements of the Hamiltonians in
\eqref{eq:H_0_neighbor} and \eqref{eq:H_int_neighbor}.  Conservatively
assuming no energetic penalty and no Pauli blocking for any inter-site
process, the dominant correction $\delta E_N$ to the interaction
energy of a lattice site with $N$ atoms and $b$ neighboring sites
(e.g.~$b=6$ in a primitive cubic lattice) is roughly bounded as
\begin{align}
  \abs{\delta E_N} \lesssim b {N\choose 2}
  \max\set{\abs{\gamma_2^{(2)}},
    \p{N-1} \abs{\gamma_{3,1}^{(2)} + \gamma_{3,2}^{(2)}}} G^2,
  \label{eq:delta_E_N}
\end{align}
where the factor of $b$ accounts for the multiplicity of neighboring
sites; the factor of ${N\choose2}$ accounts for the number of on-site
pairs of atoms which are addressed by the diagrams in
\eqref{eq:g_2_2}-\eqref{eq:g_3_2_2}; and the factor of $N-1$ on
$\gamma_{3,X}^{(2)}$ accounts for the number of atoms in a neighboring
site which are addressed by the corresponding processes.  These
factors count the number of matrix elements in the Hamiltonian with
magnitude $\sim\gamma_X^{(2)}G^2$.  The maximization in
\eqref{eq:delta_E_N} is performed because the relevant two- and
three-body processes are mutually exclusive, requiring a different
number of atoms on neighboring lattice sites.  For a conservative
bound of $\abs{\delta E_N}$, the coupling factor $G^2$ in
\eqref{eq:delta_E_N} can simply be maximized over its allowed values
for a given state of atoms on a lattice site, e.g.~$G_\g^2$ for a
state with no orbital excitations, or $\max\set{G_\g^2,G_+^2,G_-^2}$
for a state with one net orbital excitation (in both cases, assuming
no orbital excitations in neighboring sites).  In the latter case, the
bound in \eqref{eq:delta_E_N} can also be reduced by observing that to
conserve energy, it must be the excited atom which moves to a
neighboring site, which reduces the factor of ${N\choose2}$ in down to
$N-1$.  We emphasize that the bound in \eqref{eq:delta_E_N} is by no
means an exact measure of error, and is merely intended to provide a
conservative range of energies and corresponding time scales for which
inter-site effects could become relevant despite negligible
single-particle ground-state tunneling rates.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Low-excitation $M$-body Hamiltonian coefficients}
\label{sec:U_X}

The $M$-body Hamiltonians addressing low-lying orbital state
excitations can be written in the form
\begin{align}
  H_M = \sum_{\abs{\set{\mu_j}}=M}
  \p{U_{M,\g} \prod_{\alpha=1}^M \n_{\mu_\alpha,\g}
    + U_{M,+} \n_{\mu_1,\e} \prod_{\alpha=2}^M \n_{\mu_\alpha,\g}
    + U_{M,-} \c_{\mu_1,\g}^\dag \c_{\mu_2,\e}^\dag
    \c_{\mu_2,\g} \c_{\mu_1,\e} \prod_{\alpha=3}^M \n_{\mu_\alpha,\g}},
\end{align}
where the coefficients can be expanded as
$U_{M,X} = \sum_p U_{M,X}^{(p)}$ with terms $U_{M,X}^{(p)}$ at order
$p$ in the coupling constants $G_Y$.  The terms $U_{M,X}^{(p)}$ can be
determined from the $M$-body $p$-order Hamiltonians $H_M^{(p)}$
derived in section \ref{sec:low_energy}, i.e.~in \eqref{eq:H_2},
\eqref{eq:H_3_2}, \eqref{eq:H_3_3}, and \eqref{eq:H_4_3}.  For the
effective 2-, 3-, and 4-body Hamiltonians through third order in the
coupling constants, the coefficients are
\begin{align}
  U_{2,\g} = \f12 \alpha_2^{(1)} G_\g,
  &&
  U_{2,+} = \alpha_2^{(1)} G_+,
  &&
  U_{2,-} = \alpha_2^{(1)} G_-,
\end{align}
\begin{align}
  U_{3,\g}^{(2)} = - \alpha_3^{(2)} G_\g^2,
  &&
  U_{3,+}^{(2)} = - \alpha_3^{(2)} G_+ \p{G_+ + 2 G_\g},
\end{align}
\begin{align}
  U_{3,-}^{(2)} = - \alpha_3^{(2)} G_- \p{2 G_+ + G_- + 2 G_\g},
\end{align}
\begin{align}
  U_{3,\g}^{(3)}
  = \p{\alpha_3^{(3)} - \alpha_5^{(3)}} 2 G_\g^3
  - \p{\alpha_{4,3}^{(3)} + \alpha_5^{(3)}} G_\g^3,
\end{align}
\begin{align}
  U_{3,+}^{(3)}
  &= \p{\alpha_3^{(3)} - \alpha_5^{(3)}} \p{G_+^3 + 4 G_+^2 G_\g
    + G_+ G_-^2 + G_+ G_\g^2 + G_-^3 + G_-^2 G_\g} \nonumber \\
  &\quad - \p{\alpha_{4,3}^{(3)} + \alpha_5^{(3)}}
  \p{G_+^3 + G_+^2 G_\g + G_+ G_-^2 + G_+ G_\g^2 + G_-^2 G_\g},
\end{align}
\begin{align}
  U_{3,-}^{(3)}
  &= \p{\alpha_3^{(3)} - \alpha_5^{(3)}} G_- \p{3 G_+^2
    + 2 G_+ G_- + 8 G_+ G_\g + 3 G_- G_\g + G_\g^2} \nonumber \\
  &\quad - \p{\alpha_{4,3}^{(3)} + \alpha_5^{(3)}}
  G_- \p{3 G_+^2 + 2 G_+ G_- + 2 G_+ G_\g + G_-^2 + G_\g^2},
\end{align}
\begin{align}
  U_{4,\g}^{(3)}
  = \p{\alpha_{4,1}^{(3)} - \alpha_5^{(3)}} G_\g^3
  + \p{\alpha_{4,2}^{(3)} - \alpha_5^{(3)}} 2 G_\g^3,
\end{align}
\begin{align}
  U_{4,+}^{(3)}
  &= \p{\alpha_{4,1}^{(3)} - \alpha_5^{(3)}}
  2 G_+ G_\g \p{G_+ + G_\g} \nonumber \\
  &\quad + \p{\alpha_{4,2}^{(3)} - \alpha_5^{(3)}}
  G_+ \p{G_+^2 + 2 G_+ G_\g + 5 G_\g^2},
\end{align}
\begin{align}
  U_{4,-}^{(3)}
  &= \p{\alpha_{4,1}^{(3)} - \alpha_5^{(3)}}
  2 G_- G_\g \p{2 G_+ + G_- + G_\g} \nonumber \\
  &\quad + \p{\alpha_{4,2}^{(3)} - \alpha_5^{(3)}}
  G_- \p{3 G_+^2 + 3 G_+ G_- + 4 G_+ G_\g + G_-^2
    + 2 G_- G_\g + 5 G_\g^2},
\end{align}
In terms of the spatial overlap integrals defined in \eqref{eq:K_klmn}
and \eqref{eq:K}, the prefactors $\alpha_X^{(p)}$ on the coefficients
$U_X^{(p)}$ are
\begin{align}
  \alpha_2^{(1)} \equiv K,
  &&
  \alpha_3^{(2)} \equiv \sum_{n>0} \f{K_n^2}{E_n},
  &&
  \alpha_3^{(3)} \equiv \sum_{\substack{\ell+m>0\\\ell+n>0}}
  \f{K_{\ell m} K^m_n K_{\ell n}}{E_{\ell m} E_{\ell n}},
\end{align}
\begin{align}
  \alpha_{4,1}^{(3)}
  \equiv \sum_{\substack{m\ge0\\n>0}} \f{K_{mn} K_m K_n}{E_{mn} E_n},
  &&
  \alpha_{4,2}^{(3)}
  \equiv \sum_{m,n>0} \f{K_m K^m_n K_n}{E_m E_n},
  &&
  \alpha_{4,3}^{(3)}
  \equiv K \sum_{m+n>0} \f{K_{mn}^2}{E_{mn}^2},
\end{align}
\begin{align}
  \alpha_5^{(3)}
  \equiv  K \sum_{n>0} \f{K_n^2}{E_n^2}.
\end{align}

\bibliography{\jobname}

\end{document}

### 2018-07-09 olc_soc/olc_writeup.tex
added soc-related writeup and notes

\documentclass[aps,notitlepage,nofootinbib,11pt]{revtex4-1}

% linking references
\usepackage{hyperref}
\hypersetup{
  breaklinks=true,
  colorlinks=true,
  linkcolor=blue,
  filecolor=magenta,
  urlcolor=cyan,
}

%%% header / footer
\usepackage{fancyhdr} % easier header and footer management
\pagestyle{fancy} % page formatting style
\fancyhf{}
\usepackage{lastpage} % for referencing last page
\cfoot{\thepage~of \pageref{LastPage}} % "x of y" page labeling
\renewcommand{\headrulewidth}{0pt} % remove horizontal line in header

% figures
\usepackage{hyperref} % for linking references
\usepackage{graphicx,float} % for figures
\usepackage{grffile} % help latex properly identify figure extensions
\graphicspath{{./figures/}} % set path for all graphics
\usepackage[caption=false]{subfig} % subfigures ("subfloat")
\newcommand{\sref}[1]{\protect\subref{#1}}

% inline lists
\usepackage[inline]{enumitem}
\setlist[enumerate,1]{label={(\roman*)}}

%%% symbols, notations, etc.
\usepackage{physics,braket,amssymb} % physics and math packages
\usepackage{accents} % for resolving some accent (e.g. tilde) issues
\renewcommand{\t}{\text} % text in math mode
\newcommand{\f}[2]{\dfrac{#1}{#2}} % shorthand
\newcommand{\p}[1]{\left(#1\right)} % parenthesis
\renewcommand{\sp}[1]{\left[#1\right]} % square parenthesis
\renewcommand{\set}[1]{\left\{#1\right\}} % curly parenthesis
\newcommand{\bk}{\Braket} % shorthand
\renewcommand{\d}{\partial} % partial d

\usepackage{dsfont}
\newcommand{\1}{\mathds{1}}

\newcommand{\Z}{\mathbb Z}
\renewcommand{\O}{\mathcal O}

\usepackage{accents}
\newcommand{\utilde}[1]{\underaccent{\tilde}{#1}}

\begin{document}

\title{Generating new forms of spin-orbit coupling with a 1-D optical
  lattice clock}

\author{Michael A. Perlin}

\maketitle
\thispagestyle{fancy}

The primary motivation of this work is to use the ${}^{87}$Sr optical
lattice clock (OLC) to study general forms of spin-orbit coupling.  To
accomplish this task, we will draw analogies between the OLC and
trapped ions, with the hope of facilitating an exchange of ideas and
techniques between the trapped ion and optical lattice communities.
We will map degrees of freedom between the two systems as follows: the
quasi-momentum of an atom in the OLC corresponds to the on-site
orbital occupied by a trapped ion in a Coulomb lattice; the band index
of an OLC atom corresponds to a vibrational (phonon) mode of an ion;
and the (electronic) clock states of an OLC atom constitute a
pseudo-spin degree of freedom which corresponds to two optical states
of an ion.  For brevity, we will sometimes refer to the clock states
as the ``spin'' of the OLC atoms; we will not address the nuclear spin
degree of freedom of OLC atoms, and will assume that they are
nuclear-spin-polarized.

We start by considering a two level atom in a one-dimensional (1-D)
optical lattice interrogated by a plane-wave clock laser.  After a
rotating wave approximation, the dynamics of this system is
effectively described by the Hamiltonian
\begin{align}
  H
  = \f{p^2}{2m} + V_0\sin^2\p{k_L z} - \f12\delta\sigma^z
  - \f12\Omega\p{e^{ikz}\sigma^+ + e^{-ikz}\sigma^-},
  \label{eq:start}
\end{align}
where $p$ and $m$ are respectively the momentum and mass of the atom,
$V_0$ and $k_L$ are the lattice depth and wavenumber; $z$ is the
position along the lattice; $\delta=\omega-\omega_0$ is the detuning
of the clock laser frequency $\omega$ from the atomic clock transition
energy $\omega_0$; $\sigma^j$ with $j\in\set{x,y,z}$ is a Pauli matrix
addressing the spin of the atom;
$\sigma^\pm\equiv\p{\sigma^x\pm i\sigma^y}/2$ are spin raising and
lowering operators; $\Omega$ is the bare Rabi frequency of the clock
laser, and $k$ is the wavenumber of the clock laser along the lattice
axis.

\section{Quasi-momentum expansion and diagonalization}

The first two terms of the Hamiltonian in \eqref{eq:start} have
discrete translational invariance, which means that they may be
diagonalized with quasi-momentum and band index as good quantum
numbers.  The spatial eigenfunctions of these terms correspond to
solutions of the Mathieu equation, whose eigenfunctions
$\bk{z|qn}=\phi_{qn}\p{z}$ and corresponding energy eigenvalues
$E_{qn}$ are, in a periodic lattice with $L$ sites, indexed by a
quasi-momentum $q$ and band index $n$, where $q/k_L\in\Z/L$ and
$n\in\mathbb N_0$.  Additionally introducing the index
$s\in\set{\pm 1}$ to label the ground ($s=-1$) and excited ($s=1$), we
define field operators $c_{qns}$ which act on the vacuum as
$c_{qns}^\dag\ket{\t{vacuum}}=\ket{qns}$ and obey the standard
fermionic commutation relations.  The Hamiltonian in \eqref{eq:start}
can then be written in the general form
\begin{align}
  H
  = \sum_{q,n,s}\p{E_{qn}-\f12s\delta}c_{qns}^\dag c_{qns}
  - \f12\sum_{\substack{q,n,s\\g,m,r}}
  \Omega^{qns}_{gmr} c_{gmr}^\dag c_{qns},
  \label{eq:field_operators}
\end{align}
where $\Omega^{qns}_{gmr}\equiv\Omega\bk{gmr|e^{-iskz}|qns}$ is a
coupling constant between states $\ket{qns}$ and $\ket{gmr}$.  This
coupling constant vanishes unless $r=\bar s\equiv-s$.  Furthermore, in
what amounts to a stationary phase approximation (see Appendix
\ref{sec:laser_coupling}), we can say that
$\Omega^{qns}_{gm\bar s} \approx \Omega^{qns}_{gm\bar s}
\delta_{g,q-sk}$, which is a result of a momentum kick $k$ imparted
into OLC atoms when they absorb or emit a photon; this approximation
becomes exact when the clock photon momentum $k$ is commensurate with
the lattice.  Making this substitution, the Hamiltonian in
\eqref{eq:field_operators} becomes
\begin{align}
  H
  = \sum_{q,n,s}\p{E_{qn}-\f12s\delta}c_{qns}^\dag c_{qns}
  - \f12\sum_{q,n,s,m}\Omega^{qns}_{q-sk,m\bar s}
  c_{q-sk,m,\bar s}^\dag c_{qns}.
  \label{eq:kick}
\end{align}
Further redefining field operators, energies, and coupling constants
\begin{align}
  b_{qns} \equiv c_{q+sk/2,ns},
  &&
  E_{qns} \equiv E_{q+sk/2,n},
  &&
  \Omega^{qs}_{nm} = \Omega^{q+sk/2,ns}_{q-sk/2,m\bar s},
  \label{eq:transformation}
\end{align}
diagonalizes the Hamiltonian in \eqref{eq:kick} with respect to
quasi-momentum $q$:
\begin{align}
  H
  = \sum_{q,n,s}\p{E_{qns}-\f12s\delta} b_{qns}^\dag b_{qns}
  - \f12\sum_{q,n,s,m} \Omega^{qs}_{nm} b_{qm\bar s}^\dag b_{qns}.
  \label{eq:full_H}
\end{align}
This redefinition amounts to a gauge transformation of the field
operators, which shifts the OLC band structure (i.e. dispersion
relation; see Figure \ref{fig:bands}).  Appendix
\ref{sec:laser_coupling} covers some useful symmetries of
$\Omega^{qs}_{nm}$.

\begin{figure}
  \subfloat[]{
    \includegraphics[width=0.45\textwidth]{qn_bands_V10.pdf}
    \label{fig:before}
  } \subfloat[]{
    \includegraphics[width=0.45\textwidth]{qns_bands_V10.pdf}
    \label{fig:after}
  }
  \caption{Band diagrams of the Sr-87 OLC with a lattice depth
    $V_0=10E_R$ (where $E_R=k_L^2/2m\approx22~\t{kHz}$ is the lattice
    recoil energy) \sref{fig:before} before, and \sref{fig:after}
    after the gauge transformation in \eqref{eq:transformation}. The
    legend in \sref{fig:after} indicates the clock state (spin) for
    each line.}
  \label{fig:bands}
\end{figure}

A few of observations about the Hamiltonian in \eqref{eq:full_H}.
First, there is no coupling between states with different
(gauge-transformed) quasi-momenta $q$.  Once initialized, therefore,
the single-particle physics of the this OLC is reducible to the
dynamics within each sub-manifold of states with fixed quasi-momentum
$q$.  This fact is in analogy with trapped ions which are pinned to a
single site of a Coulomb crystal.  Second, there is no direct coupling
between different states with the same spin $s$; such coupling only
occurs at second order in the coupling strength $\Omega$.  In
principle, it is possible to introduce first-order coupling between
states with different band index $n$ and the same spin $s$ by
modulating the lattice in either phase or amplitude, but for now we
leave this consideration for future work.  Third, the Hamiltonian in
\eqref{eq:full_H} exhibits the standard type of spin-orbit coupling
investigated in the $^{87}$Sr OLC at JILA, in which different
quasi-momenta $q$ are energetically favorable for different spins $s$.
This spin-orbit coupling is similar to that encountered in a condensed
matter setting, and can lead to interesting topological physics.
Finally, the 1-D OLC Hamiltonian exhibits a spin-phonon-like coupling
between spin $s$ and band index $n$, which we will see more clearly in
Section \ref{sec:ions}.

\section{Mapping onto trapped ions}
\label{sec:ions}

In this section, we illustrate more clearly the analogy between the
1-D optical lattice clock and a 1-D crystal of trapped ions.  To
simplify our analysis, assume single-particle occupation each
quasi-momentum in the OLC, which translates to having a single ion per
site in a Coulomb crystal.  We will also only consider transitions
between adjacent bands (i.e. vibrational modes). We now define the
``lowering operators'' $a_{qnm}\equiv\sum_s\op{q,n-1,s}{qns}$; in
addition to
\begin{align}
  \bar E_{qn} \equiv \f12\p{E_{qn,+} + E_{qn,-}},
  &&
  \Delta_{qn} \equiv E_{qn,+} - E_{qn,-},
\end{align}
which are the mean energy ($\bar E_{qn}$) and energy splitting
($\Delta_{qn}$) between spin states within one band (i.e. between
$\ket{qn,+}$ and $\ket{qn,-}$); and finally
\begin{align}
  \Omega_{qn} \equiv \Omega^{q,+}_{nn} = \Omega^{q,-}_{nn},
  &&
  \Lambda_{qn}^x
  \equiv \f12\p{\Omega^{q,+}_{n-1,n} + \Omega^{q,-}_{n-1,n}},
  &&
  \Lambda_{qn}^y
  \equiv \f12\p{\Omega^{q,+}_{n-1,n} - \Omega^{q,-}_{n-1,n}},
\end{align}
which are intra-band ($\Omega_{qn}$) and inter-band ($\Lambda_{qn}^x$,
$\Lambda_{qn}^y$) coupling coefficients, respectively corresponding to
coefficients for carrier and side-band transitions.  In terms of these
quantities, the Hamiltonian from \eqref{eq:full_H} can be written as
\begin{align}
  H
  = \sum_{q,n} \sp{\op{qn}
    \p{\bar E_{qn} + \sp{\Delta_{qn}-\delta}S^z - \Omega_{qn}S^x}
    - \Lambda_{qn}^x \p{a_{qn}^\dag + a_{qn}}S^x
    - \Lambda_{qn}^y i\p{a_{qn}^\dag - a_{qn}}S^y},
  \label{eq:separated}
\end{align}
where we have introduced the on-band spin operators
$S^j\equiv\sigma^j/2$. The full derivation of the Hamiltonian in
\eqref{eq:separated} from that in \eqref{eq:full_H} is provided in
Appendix \ref{sec:spin_separation}.

The operators $a_{qn}+a_{qn}^\dag$ and $i\p{a_{qn}^\dag-a_{qn}}$ in
\eqref{eq:separated} can be identified with position and momentum
operators in a trapped ion Hamiltonian. In this sense we are
engineering a second type of spin-orbit coupling in the OLC: between
spin and band modes, which is analogous to coupling between spin and
vibrational phonon modes in a trapped ion system.

\subsection{Deep lattice approximation}

As the lattice depth $V_0$ is increased relative to the lattice recoil
energy $E_R\equiv k_L^2/2m$, i.e. when $V_0\gg E_R$, the energies
$E_{qns}$ and coupling coefficients $\Omega^{qs}_{nm}$ lose their
dependence on quasi-momentum $q$, which implies that
\begin{align}
  \bar E_{qn} \to E_n \equiv \f1L\sum_q E_{qn},
  &&
  \Delta_{qn} \to 0,
\end{align}
\begin{align}
  \Omega_{qn} \to \Omega_n \equiv \f1L\sum_q \Omega_{qn},
  &&
  \Lambda_{qn}^x \to 0,
  &&
  \Lambda_{qn}^y \to
  \Lambda_n^y \equiv \f1L\sum_q \Lambda_{qn}^y,
\end{align}
where $L$ is the number of lattice sites in the OLC. The fact that
$\Lambda_{qn}^x$ vanishes while $\Lambda_{qn}^y$ is preserved is a
consequence of the symmetries of $\Omega^{qs}_{nm}$, which we derive
in Appendix \ref{sec:laser_coupling} and summarize in
\eqref{eq:symmetries}. With these simplifications, the single-particle
Hamiltonian in \eqref{eq:separated} reduces to the simpler form
\begin{align}
  H
  = \sum_{q,n}\sp{\op{qn}\p{E_n - \delta S^z - \Omega_n S^x}
    - \Lambda_n^y i\p{a_{qn}^\dag - a_{qn}}S^y}.
  \label{eq:deep}
\end{align}
This Hamiltonian has lost the standard type of spin-orbit coupling
investigated in the 1-D OLC, in that there is no dependence of any
dynamical variables on the crystal momentum $q$.  Nonetheless, the
spin-phonon like coupling is preserved through the
$a_{qn}^\dag-a_{qn}$ term.

\section{Dynamics}
\label{sec:dynamics}

We now consider the OLC Hamiltonian after diagonalization in
quasi-momentum, i.e. \eqref{eq:full_H}, in the interaction picture of
the bare energies $E_{qns}-s\delta/2$:
\begin{align}
  H_I
  = -\f12\sum_{q,n,s,m} \Omega^{qs}_{nm}
  \exp\sp{i\p{\Delta^{qs}_{nm}+s\delta}t}
  b_{qm\bar s}^\dag b_{qns},
  \label{eq:full_H_I}
\end{align}
where $\Delta^{qs}_{nm}\equiv E_{qm\bar s}-E_{qns}$ is the energy gap
between the single-particle states $\ket{qm\bar s}$ and $\ket{qns}$ at
no detuning ($\delta=0$). With the ${}^{87}$Sr OLC, in practice the
Rabi frequency $\Omega$ is much smaller than the energy difference
between two bands, such that
\begin{align}
  \abs{\Omega^{qs}_{nm}} < \Omega\ll\abs{\Delta^{qs}_{nm}}
  \label{eq:weak_laser}
\end{align}
for any $q,n,s$ and $m\ne n$. Without detuning, all inter-band
couplings in \eqref{eq:full_H_I} therefore vanish by the secular
approximation. In the trapped ion system, \eqref{eq:weak_laser}
corresponds to having an interrogation laser which is too weak to
excite vibrational modes of the ions. While it is possible to set a
static detuning $\delta$ which bridges the energy difference between
different bands, doing so will still ultimately result in dynamics
which are reducible to that of a two-level system. In order to induce
nontrivial multi-band dynamics, we must therefore introduce time
dependence into the Hamiltonian. A simple means of introducing time
dependence is to modulate the detuning $\delta$ or Rabi frequency
$\Omega$, which respectively correspond to frequency and amplitude
modulation of the clock-state interrogation laser.

\subsection{Frequency modulation}
\label{sec:freq_mod}

Modulating the detuning as
$\delta\p{t}=\delta_0-\tilde\delta\cos\p{\nu t}$ for some mean
$\delta_0$, modulation amplitude $\tilde\delta$, and modulation
frequency $\nu$ results in the Hamiltonian
\begin{align}
  H_I
  = -\f12\sum_{\substack{q,n,s\\m,\kappa}}
  J\p{\kappa,s\tilde\delta/\nu} \Omega^{qs}_{nm}
  \exp\sp{i\p{\Delta^{qs}_{nm}+s\delta_0-\kappa\nu}t}
  b_{qm\bar s}^\dag b_{qns},
  \label{eq:freq_mod_H}
\end{align}
where $\kappa\in\Z$ and $J\p{n,x}$ is the $n$-th order Bessel function
of the first kind evaluated at $x$. If $\nu\gg\Omega$ and the OLC is
initialized in the state $\ket\phi=\ket{q_0n_0s_0}$, then the dynamics
induced by \eqref{eq:freq_mod_H} can be equivalently realized by the
effective time-independent Hamiltonian
\begin{align}
  H_I^{\t{eff}}
  = \sum_{q,n,s} \epsilon_{qns}^\phi b_{qns}^\dag b_{qns}
  - \f12\sum_{q,n,s,m} J\p{\kappa^{qs}_{nm},s\tilde\delta/\nu}
  \Omega^{qs}_{nm} b_{qm\bar s}^\dag b_{qns},
  \label{eq:freq_mod_H_eff}
\end{align}
where $\kappa^{qs}_{nm}\in\Z$ minimizes
$\abs{\Delta^{qs}_{nm}+s\delta_0-\kappa^{qs}_{nm}\nu}$ and the reduced
energies $\epsilon_{qns}^\phi$ satisfy
\begin{align}
  E_{qns} - \f12s\delta_0
  = E_{q_0n_0s_0} - \f12s_0\delta_0
  + \ell_{qns}^\phi\nu + \epsilon_{qns}^\phi
  \label{eq:reduced_E}
\end{align}
with $\ell_{qns}^\phi\in\Z$ and $\abs{\epsilon_{qns}^\phi}<\nu/2$. The
derivations of \eqref{eq:freq_mod_H} and \eqref{eq:freq_mod_H_eff} are
provided in Appendix \ref{sec:freq_mod_derivation}.

\begin{figure}[hp]
  \captionsetup[subfloat]{farskip=1pt,captionskip=1pt}
  \subfloat[First excited band, $V_0=80E_R$]{
    \includegraphics[width=0.45\textwidth]{freq_mod_band_V80.pdf}
    \label{fig:freq_mod_band_V80}
  } \subfloat[Excited clock state, $V_0=80E_R$]{
    \includegraphics[width=0.45\textwidth]{freq_mod_spin_V80.pdf}
    \label{fig:freq_mod_spin_V80}
  }
  \\
  \subfloat[First excited band, $V_0=40E_R$]{
    \includegraphics[width=0.45\textwidth]{freq_mod_band_V40.pdf}
    \label{fig:freq_mod_band_V40}
  } \subfloat[Excited clock state, $V_0=40E_R$]{
    \includegraphics[width=0.45\textwidth]
    {freq_mod_spin_V40.pdf}
    \label{fig:freq_mod_spin_V40}
  }
  \\
  \subfloat[First excited band, $V_0=10E_R$]{
    \includegraphics[width=0.45\textwidth]
    {freq_mod_band_V10.pdf}
    \label{fig:freq_mod_band_V10}
  } \subfloat[Excited clock state, $V_0=10E_R$]{
    \includegraphics[width=0.45\textwidth]
    {freq_mod_spin_V10.pdf}
    \label{fig:freq_mod_spin_V10}
  }
  \caption{Time evolution of states initially in the lowest band
    ($n=0$) and ground clock state ($s=-1$) subject to the Hamiltonian
    in \eqref{eq:freq_mod_H_eff} with a detuning
    $\delta=\Delta\cos\p{\Delta t}$ for $\Delta$ equal to the mean
    energy gap between the lowest two bands.  Color indicates the
    population of the state specified in the captions, which also
    indicate the lattice depths.  The interaction strength
    $\Omega^{0,-}_{1,0}/2\pi$ is approximately $43~\t{Hz}$ in a deep
    lattice with $V_0=80E_R$, $51~\t{Hz}$ in a medium lattice with
    $V_0=40E_R$, and $58~\t{Hz}$ in a shallow lattice with
    $V_0=40E_R$.  In all cases, $\Omega=1~\t{kHz}$.}
  \label{fig:freq_mod}
\end{figure}

At zero mean detuning ($\delta_0=0$), any choice of
$\nu=\Delta^{qs}_{nm}/\ell$ for $\ell\in\Z$ can, in principle, induce
multi-band dynamics.  As an example, Figure \ref{fig:freq_mod} shows
the time evolution of atoms initially in the lowest band and ground
clock state (i.e. $\ket{q,0,-}$) subject to a detuning
$\delta\p{t}=\Delta\cos\p{\Delta t}$, where $\Delta$ is the mean
energy gap between the lowest two bands (vibrational modes).  The
plots in this figure show the expectation values
$\bk{\op{q}\otimes\O_B\otimes\O_S}$ over time for
$\O_B,\O_S\in\set{\1,\op{1}}$ and a variety of lattice depths $V_0$.
Here $\O_{B,S}=\1$ traces out over the corresponding degree of freedom
(i.e. band or spin), while $\O_{B,S}=\op{1}$ appropriately selects out
either the first excited band or the excited clock state.  The value
of $\bk{\op{q}\otimes\op{1}\otimes\1}$ thus measures the total
population of atoms in the first excited band (independent of clock
state), while $\bk{\op{q}\otimes\1\otimes\op{1}}$ measures the total
population of atoms in the excited clock state (independent of band).

In a deep lattice ($V_0=80E_R$), the $q$-dependence of the energy gaps
$\Delta^{qs}_{nm}$ vanishes. We therefore see no spin-orbit coupling
in a deep lattice (Figures \ref{fig:freq_mod_band_V80} and
\ref{fig:freq_mod_spin_V80}), as the modulation frequency $\nu$ is
resonant with these gaps for all $q$. In the medium depth lattice
($V_0=40E_R$; Figures \ref{fig:freq_mod_band_V40} and
\ref{fig:freq_mod_spin_V40}), $\nu$ is only approximately resonant
with the energy gaps for all $q$, leading to some spin-orbit
coupling. Finally, spin-orbit coupling is most prominent in the
shallow lattice ($V_0=10E_R$; Figures \ref{fig:freq_mod_band_V10} and
\ref{fig:freq_mod_spin_V10}), in which the energy gaps, and thus OLC
dynamics, are strongly selective on quasi-momentum. In all cases,
there is evidence of spin-phonon-like coupling in the form of common
features between the excited band and excited clock state populations.

\subsection{Amplitude modulation}
\label{sec:amp_mod}

We can alternately modulate the amplitude of the clock laser as
$\Omega\to\Omega\cos\p{\nu t}$ for some frequency $\nu$, which results
in the Hamiltonian
\begin{align}
  H_I
  = -\f14\sum_{\substack{q,n,s\\m,r}} \Omega^{qs}_{nm}
  \exp\sp{i\p{\Delta^{qs}_{nm}+s\delta+r\nu}t}
  b_{qm\bar s}^\dag b_{qns},
  \label{eq:amp_mod_H}
\end{align}
for $r\in\set{-1,1}$. At no detuning, the amplitude modulation
frequency $\nu$ can be chosen to bridge some particular energy gap
$\Delta^{qs}_{nm}$ to induce transitions between different
bands. Similarly to the restriction on dynamics induced by the
time-independent OLC Hamiltonian in \eqref{eq:full_H_I} due to the
limitation in \eqref{eq:weak_laser}, however, in the ${}^{87}$Sr OLC
the differences between the different band gaps $\Delta^{qs}_{nm}$ are
generally larger than the Rabi coupling $\Omega$. The dynamics of the
OLC with amplitude modulation of the clock laser at no detuning thus
generally reduces to that of a two-level system.

While most choices of detuning $\delta$ do not change the above
argument, the symmetric effect of $\delta$ on different clock states
means that it is possible to engineer four-level dynamics in the OLC
for quasi-momenta $q$ with
$\Delta^{q,-}_{nm}\approx\Delta^{q,+}_{nm}$. Letting
\begin{align}
  \utilde\Delta^{qs}_{nm}
  \equiv \Delta^{qs}_{nm} + s\delta
  = \p{E_{qm\bar s} - \f12\bar s\delta} - \p{E_{qns} - \f12s\delta}
\end{align}
be the total band gap between $\ket{qm\bar s}$ and $\ket{qns}$ at
detuning $\delta$, if $\Delta^{0,-}_{nm}=\Delta^{0,+}_{nm}$ and we set
$\delta\approx\Delta^{0,\pm}_{1,0}/2$, for example, then one can work
out that
\begin{align}
  \utilde\Delta^{0,+}_{0,0}
  = \utilde\Delta^{0,-}_{1,0}
  = \utilde\Delta^{0,+}_{1,1}
  = \Delta^{0,\pm}_{1,0}/2
  \approx \delta
  \label{eq:gaps}
\end{align}
Setting $\nu=\delta$ then simultaneously couples
$\ket{0,0,-}\leftrightarrow\ket{0,0,+}
\leftrightarrow\ket{0,1,-}\leftrightarrow\ket{0,1,+}$, while coupling
to all higher bands can be neglected by the secular approximation.

\begin{figure}
  \captionsetup[subfloat]{farskip=1pt,captionskip=1pt}
  \subfloat[$V_0=10E_R$]{
    \includegraphics[width=0.45\textwidth]{qns_bands_detuned_V10.pdf}
    \label{fig:test}
  } \subfloat[$V_0=40E_R$]{
    \includegraphics[width=0.45\textwidth]{qns_bands_detuned_V40.pdf}
    \label{fig:test2}
  }
  \caption{Band diagrams of the Sr-87 OLC with a detuning
    $\delta=\Delta/2$ for two lattice depths $V_0$.  Legends indicate
    the clock state (spin) for each line.}
  \label{fig:bands_detuned}
\end{figure}

Figure \ref{fig:bands_detuned} shows the band diagrams which results
from setting $\delta=\Delta/2$ with lattice depths of $V_0=10E_R$ and
$V_0=40E_R$. While the energy gaps between the first four energy
levels (corresponding to both clock states in the lowest two bands)
are identical (at $q=0$), the gap to the fifth level (in the third
band) is different by an amount sufficiently large to make the fifth
level off-resonant for state transition.

Figure \ref{fig:amp_mod} shows the time evolution of atoms initially
in the lowest band and ground clock state (i.e. $\ket{q,0,-}$) subject
to a static detuning $\delta=\Delta/2$ and amplitude modulation with
frequency $\nu=\Delta/2$. As before, spin-orbit coupling is most
prominent in a shallow lattice, and is suppressed in a deep lattice;
spin-phonon-like coupling is present for all lattice depths.

Note that similarly to the case in Section \ref{sec:freq_mod}, if
$\nu\gg\Omega$ and the OLC is initialized in the state
$\ket\phi=\ket{q_0n_0s_0}$, then dynamics induced by
\eqref{eq:amp_mod_H} may be equivalently realized by the effective
time-independent Hamiltonian
\begin{align}
  H_I^{\t{eff}}
  = \sum_{q,n,s} \epsilon_{qns}^\phi b_{qns}^\dag b_{qns}
  - \f14\sum_{q,n,s,m} C\p{\kappa^{qs}_{nm}} \Omega^{qs}_{nm}
  b_{qm\bar s}^\dag b_{qns}
\end{align}
where $\kappa^{qs}_{nm}\in\Z$ minimizes
$\abs{\Delta^{qs}_{nm}+s\delta_0-\kappa^{qs}_{nm}\nu}$;
$C\p{\kappa}\equiv1$ when $\abs{\kappa}=1$ and $C\p{\kappa}\equiv0$
otherwise; and the reduced energies $\epsilon_{qns}^\phi$ satisfy
\eqref{eq:reduced_E} with $\ell_{qns}^\phi\in\Z$ and
$\abs{\epsilon_{qns}^\phi}<\nu/2$.

\begin{figure}[h!]
  \captionsetup[subfloat]{farskip=1pt,captionskip=1pt}
  \subfloat[First excited band, $V_0=80E_R$]{
    \includegraphics[width=0.45\textwidth]{amp_mod_band_V80.pdf}
    \label{fig:amp_mod_band_V80}
  } \subfloat[Excited clock state, $V_0=80E_R$]{
    \includegraphics[width=0.45\textwidth]{amp_mod_spin_V80.pdf}
    \label{fig:amp_mod_spin_V80}
  }
  \\
  \subfloat[First excited band, $V_0=40E_R$]{
    \includegraphics[width=0.45\textwidth]{amp_mod_band_V40.pdf}
    \label{fig:amp_mod_band_V40}
  } \subfloat[Excited clock state, $V_0=40E_R$]{
    \includegraphics[width=0.45\textwidth]{amp_mod_spin_V40.pdf}
    \label{fig:amp_mod_spin_V40}
  }
  \\
  \subfloat[First excited band, $V_0=10E_R$]{
    \includegraphics[width=0.45\textwidth]{amp_mod_band_V10.pdf}
    \label{fig:amp_mod_band_V10}
  } \subfloat[Excited clock state, $V_0=10E_R$]{
    \includegraphics[width=0.45\textwidth]{amp_mod_spin_V10.pdf}
    \label{fig:amp_mod_spin_V10}
  }
  \caption{Plots corresponding to those in Figure \ref{fig:freq_mod},
    but for a static detuning $\delta=-\Delta/2$ and amplitude
    modulated clock laser which takes
    $\Omega\to\Omega\cos\p{\Delta t}$.}
  \label{fig:amp_mod}
\end{figure}

\newpage
\appendix

\section{Properties of the laser-induced coupling constants}
\label{sec:laser_coupling}

Spatial eigenfunctions of atoms on a 1-D periodic lattice are indexed
by quasi-momentum $q$ and band $n$, and take the form
\begin{align}
  \bk{z|qn}
  = \phi_{qn}\p{z}
  = e^{iqz} u_{qn}\p{z}
  = e^{iqz} \sum_{\kappa=-\infty}^\infty c_{qn}^{(\kappa)} e^{i2k_L z\kappa},
\end{align}
where $k_L$ is the wavenumber of the lattice photons, and all
coefficients $c_{qn}^{(\kappa)}\in\mathbb R$.  We can thus expand
\begin{align}
  \Omega^{qns}_{gm}
  \equiv \Omega\bk{gm|e^{-iskz}|qn}
  = \Omega \int dz~ e^{i\p{q-sk-g}z}
  \sum_{\kappa,\ell} e^{i2k_Lz\p{\kappa-\ell}}
  c_{gm}^{(\ell)} c_{qn}^{(\kappa)},
\end{align}
which vanishes unless $g\approx q-sk$\footnote{This claim is formally
  a stationary phase approximation, and becomes exact if $k$ is
  commensurate with the lattice.} and $\kappa=\ell$, so
$\Omega^{qns}_{gm}\approx\Omega^{qns}_{q-sk,m}\delta_{g,q-sk}$ only
couples the state $\ket{qns}$ to $\ket{q-sk,m\bar s}$.  Defining
\begin{align}
  \Omega^{qs}_{nm}
  \equiv \Omega^{q+sk/2,ns}_{q-sk/2,m}
  = \Omega \sum_\kappa c_{q-sk/2,m}^{(\kappa)} c_{q+sk/2,n}^{(\kappa)},
\end{align}
we can say that
\begin{align}
  \Omega^{qs}_{nm}
  = \Omega^{q+sk/2,ns}_{q-sk/2,m}
  = \p{\Omega^{q+sk/2,ns}_{q-sk/2,m}}^*
  = \Omega^{q-sk/2,m\bar s}_{q+sk/2,n}
  = \Omega^{q+\bar sk/2,m\bar s}_{q-\bar sk/2,n}
  = \Omega^{q\bar s}_{mn}.
  \label{eq:flip_both}
\end{align}
We can also use the fact that
$\phi_{qn}\p{z}=\p{-1}^n\phi_{-q,n}\p{z}^*$ to say
\begin{align}
  \Omega^{qs}_{nm}
  = \Omega\int dz~ \phi_{q-sk/2,m}\p{z}^* \phi_{q+sk/2,n}\p{z}
  = \p{-1}^{n+m}\Omega^{-q,s}_{mn}.
  \label{eq:flip_bands}
\end{align}
To summarize \eqref{eq:flip_both} and \eqref{eq:flip_bands}:
\begin{align}
  \Omega^{qs}_{nm}
  = \p{-1}^{n+m} \Omega^{-q,s}_{mn}
  = \p{-1}^{n+m} \Omega^{-q,\bar s}_{nm}
  = \Omega^{q\bar s}_{mn}.
  \label{eq:symmetries}
\end{align}

\section{Separating out the pseudo-spin degree of freedom}
\label{sec:spin_separation}

We start with the 1-D OLC Hamiltonian after diagonalization in
quasi-momentum, i.e.
\begin{align}
  H
  = \sum_{q,n,s}\p{E_{qns}-\f12s\delta} b_{qns}^\dag b_{qns}
  - \f12\sum_{q,n,s,m} \Omega^{qs}_{nm} b_{qm\bar s}^\dag b_{qns},
\end{align}
and define
\begin{align}
  \bar E_{qn} \equiv \f12\p{E_{qn,+} + E_{qn,-}},
  &&
  \Delta_{qn} \equiv E_{qn,+} - E_{qn,-},
\end{align}
\begin{align}
  \Omega_{qn} \equiv \Omega^{q,+}_{nn} = \Omega^{q,-}_{nn},
  &&
  \Lambda^{qs}_{nm}
  \equiv \Omega^{qs}_{n,n-m} = \Omega^{q\bar s}_{n-m,n},
\end{align}
in terms of which
\begin{multline}
  H = \sum_{q,n,s}\sp{\p{\bar E_{qn} + \f12 s \sp{\Delta_{qn}-\delta}}
    b_{qns}^\dag b_{qns}
    - \f12\Omega_{qn} b_{qn\bar s}^\dag b_{qns}} \\
  - \f12\sum_{\substack{q,n,s\\0<m\le n}}\Lambda^{qs}_{nm}
  \p{b_{q,n-m,\bar s}^\dag b_{qns} + b_{qns}^\dag b_{q,n-m,\bar s}}.
  \label{eq:lambda}
\end{multline}
In order to separate out the spin degree of freedom, we consider only
at most single-particle occupation of each quasi-momentum, and define
the ``lowering operators'' $a_{qnm}\equiv\op{q,n-m}{qn}$ in addition
to the on-band spin operators $S^j\equiv\sigma^j/2$ which act only on
the spin degree of freedom.  We can rewrite \eqref{eq:lambda} in terms
of these operators for single particles as
\begin{align}
  H
  = \sum_{q,n} \op{qn}\sp{\bar E_{qn}
    + \p{\Delta_{qn}-\delta} S^z - \Omega_{qn} S^x}
  - \sum_{\substack{q,n\\0<m\le n}}\Lambda^{qs}_{nm}
  \p{a_{qnm} S^{\bar s} + a_{qnm}^\dag S^s}.
\end{align}
We now expand
\begin{multline}
  \sum_s\Lambda^{qs}_{nm}\p{a_{qnm} S^{\bar s} + a_{qnm}^\dag S^s} =
  \Lambda^{q,+}_{nm}\p{a_{qnm}S^- + a_{qnm}^\dag S^+}
  + \Lambda^{q,-}_{nm}\p{a_{qnm}S^+ + a_{qnm}^\dag S^-} \\
  = \f12S^x \p{\Lambda^{q,+}_{nm} + \Lambda^{q,-}_{nm}} \p{a_{qnm} +
    a_{qnm}^\dag} + \f12 iS^y \p{\Lambda^{q,+}_{nm} -
    \Lambda^{q,-}_{nm}} \p{-a_{qnm} + a_{qnm}^\dag},
\end{multline}
which motivates the definitions
\begin{align}
  \Lambda_{qnm}^x
  \equiv \f12\p{\Lambda^{q,+}_{nm} + \Lambda^{q,-}_{nm}}
  = \f12\p{\Omega^{q,+}_{n-m,n} + \Omega^{q,-}_{n-m,n}}
  = \f12\p{\Omega^{q,+}_{n-m,n} + \p{-1}^m\Omega^{-q,+}_{n-m,n}}, \\
  \Lambda_{qnm}^y
  \equiv \f12\p{\Lambda^{q,+}_{nm} - \Lambda^{q,-}_{nm}}
  = \f12\p{\Omega^{q,+}_{n-m,n} - \Omega^{q,-}_{n-m,n}}
  = \f12\p{\Omega^{q,+}_{n-m,n} - \p{-1}^m\Omega^{-q,+}_{n-m,n}},
\end{align}
such that
\begin{multline}
  H = \sum_{q,n}\op{qn}
  \sp{\bar E_{qn} + \p{\Delta_{qn}-\delta}S^z - \Omega_{qn}S^x} \\
  - \sum_{\substack{q,n\\m>0}}
  \sp{\Lambda_{qnm}^x \p{a_{qnm}^\dag + a_{qnm}}S^x
    + \Lambda_{qnm}^y i\p{a_{qnm}^\dag - a_{qnm}}S^y}.
\end{multline}
Letting $a_{qn}\equiv a_{qn,1}$;
$\Lambda_{qn}^j\equiv\Lambda_{qn,1}^j$ for $j=x,y$; and neglecting all
inter-band (side-band) couplings beyond nearest bands, we arrive at
the Hamiltonian given in \eqref{eq:separated}:
\begin{align}
  H
  = \sum_{q,n} \sp{\op{qn}
    \p{\bar E_{qn} + \sp{\Delta_{qn}-\delta}S^z - \Omega_{qn}S^x}
    - \Lambda_{qn}^x \p{a_{qn}^\dag + a_{qn}}S^x
    - \Lambda_{qn}^y i\p{a_{qn}^\dag - a_{qn}}S^y}.
\end{align}

\section{Frequency modulation of the clock laser}
\label{sec:freq_mod_derivation}

In the interaction picture of the single-particle state energies
$E_{qns}-s\delta/2$, the Hamiltonian of the 1-D OLC is
\begin{align}
  H_I
  = - \f12\sum_{q,n,s,m} \Omega^{qs}_{nm}
  \exp\sp{i\p{\Delta^{qs}_{nm}+s\delta}t} b_{qm\bar s}^\dag b_{qns},
\end{align}
where $\Delta^{qs}_{nm}\equiv E_{qm\bar s}-E_{qns}$ is the energy gap
between the single-particle states $\ket{qm\bar s}$ and $\ket{qns}$ at
no detuning. If we modulate the detuning as
$\delta\p{t}=\delta_0-\tilde\delta\cos\p{\nu t}$ for some mean
$\delta_0$, modulation amplitude $\tilde\delta$, and modulation
frequency $\nu$, then the Hamiltonian becomes
\begin{align}
  H_I
  = - \f12\sum_{q,n,s,m} \Omega^{qs}_{nm}
  \exp\sp{i\utilde\Delta^{qs}_{nm}t-is\tilde\delta\sin\p{\nu t}/\nu}
  b_{qm\bar s}^\dag b_{qns},
\end{align}
where
\begin{align}
  \utilde\Delta^{qs}_{nm}
  \equiv \Delta^{qs}_{nm} + s\delta_0
  = \p{E_{qm\bar s}-\f12\bar s\delta_0} - \p{E_{qns}-\f12s\delta_0}
  = \utilde E_{qm\bar s} - \utilde E_{qns}.
\end{align}
is the total band gap between $\ket{qm\bar s}$ and $\ket{qns}$ at the
mean detuning $\delta_0$ and
$\utilde E_{qns}\equiv E_{qns}-\f12s\delta_0$.

Letting $\tau\equiv\nu t$ and $\beta\equiv\tilde\delta/\nu$, the
exponential $\exp\p{is\beta\sin\tau}$ is periodic in $\tau$ with
period $2\pi$, which means that we can expand it in a Fourier series
as
\begin{align}
  e^{-is\beta\sin\tau}
  = \sum_{\kappa=-\infty}^\infty e^{-i\kappa\tau}\f1{2\pi}
  \int_{-\pi}^\pi dx~ e^{i\kappa x - is\beta\sin x}
  = \sum_\kappa J\p{\kappa,s\beta} e^{-i\kappa\tau}
\end{align}
where $J\p{n,x}$ is the $n$-th order Bessel function of the first kind
evaluated at $x$. It follows that
\begin{align}
  H_I
  = -\f12\sum_{\substack{q,n,s\\m,\kappa}}
  J\p{\kappa,s\tilde\delta/\nu} \Omega^{qs}_{nm}
  \exp\sp{i\p{\utilde\Delta^{qs}_{nm}-\kappa\nu}t}
  b_{qm\bar s}^\dag b_{qns},
\end{align}
which is precisely the Hamiltonian given in
\eqref{eq:freq_mod_H}. Letting $\kappa^{qs}_{nm}\in\Z$ minimize
$\abs{\utilde\Delta^{qs}_{nm}-\kappa^{qs}_{nm}\nu}$ and assuming that
$\nu\gg\Omega$, by the secular approximation
\begin{align}
  H_I
  = -\f12\sum_{q,n,s,m}
  J\p{\kappa^{qs}_{nm},s\tilde\delta/\nu} \Omega^{qs}_{nm}
  \exp\sp{i\p{\utilde\Delta^{qs}_{nm}-\kappa^{qs}_{nm}\nu}t}
  b_{qm\bar s}^\dag b_{qns}.
  \label{eq:freq_mod_H_derived}
\end{align}
If an atom in the OLC is initially in the state
$\ket\phi=\ket{q_0n_0s_0}$, then the only relevant states for all OLC
dynamics will be those for which
\begin{align}
  \utilde E_{qns}
  = \utilde E_{q_0n_0s_0} + \ell_{qns}^\phi\nu + \epsilon_{qns}^\phi
  \label{eq:relevant_condition}
\end{align}
with $\ell_{qns}^\phi\in\Z$ and reduced energies $\epsilon_{qns}^\phi$
satisfying $\abs{\epsilon_{qns}^\phi}\lesssim\Omega\ll\nu$. For all
relevant states, therefore,
\begin{align}
  \utilde\Delta^{qs}_{nm} - \kappa^{qs}_{nm}\nu
  = \utilde E_{qm\bar s} - \utilde E_{qns} - \kappa^{qs}_{nm}\nu
  = \p{\ell_{qm\bar s}^\phi - \ell_{qns}^\phi - \kappa^{qs}_{nm}}\nu
  + \epsilon_{qm\bar s}^\phi - \epsilon_{qns}^\phi.
\end{align}
As $\abs{\epsilon_{qm\bar s}^\phi-\epsilon_{qns}^\phi}<\nu/2$ for all
relevant $q,n,s,m$, minimizing
$\abs{\utilde\Delta^{qs}_{nm}-\kappa^{qs}_{nm}\nu}$ forces
$\kappa^{qs}_{nm}=\ell_{qm\bar s}^\phi-\ell_{qns}^\phi$. The
Hamiltonian in \eqref{eq:freq_mod_H_derived} is then
\begin{align}
  H_I
  = -\f12\sum_{q,n,s,m}
  J\p{\kappa^{qs}_{nm},s\tilde\delta/\nu} \Omega^{qs}_{nm}
  \exp\sp{i\p{\epsilon_{qm\bar s}^\phi-\epsilon_{qns}^\phi}t}
  b_{qm\bar s}^\dag b_{qns},
\end{align}
and, with appropriate choice of interaction picture, is equivalent to
the effective Hamiltonian
\begin{align}
  H_I^{\t{eff}}
  = \sum_{q,n,s} \epsilon_{qns}^\phi b_{qns}^\dag b_{qns}
  - \f12\sum_{q,n,s,m} J\p{\kappa^{qs}_{nm},s\tilde\delta/\nu}
  \Omega^{qs}_{nm} b_{qns}^\dag b_{qm\bar s},
  \label{eq:freq_mod_H_eff_derived}
\end{align}
which has the great advantage over \eqref{eq:freq_mod_H_derived} of
lacking time dependence. In using \eqref{eq:freq_mod_H_eff_derived}
for e.g. computing a propagator or otherwise simulating OLC dynamics,
there is no need to actually keep track of which states are
``relevant'', as all irrelevant states will generally have reduced
energies $\epsilon_{qns}^\phi\sim\nu/2\gg\Omega$, and therefore
automatically decouple from all relevant states. A similar trick can
be used to effectively simulate OLC dynamics with amplitude modulation
as in \eqref{eq:amp_mod_H}, which results in the Hamiltonian
\begin{align}
  H_I^{\t{eff}}
  = \sum_{q,n,s} \epsilon_{qns}^\phi b_{qns}^\dag b_{qns}
  - \f14\sum_{q,n,s,m} C\p{\kappa^{qs}_{nm}} \Omega^{qs}_{nm}
  b_{qm\bar s}^\dag b_{qns},
\end{align}
for $C\p{\kappa}\equiv1$ when $\abs{\kappa}=1$ and
$C\p{\kappa}\equiv0$ otherwise.

\end{document}

### 2018-07-09 olc_soc/soc_interactions.tex
added soc-related writeup and notes

\documentclass[aps,notitlepage,nofootinbib,11pt]{revtex4-1}

% linking references
\usepackage{hyperref}
\hypersetup{
  breaklinks=true,
  colorlinks=true,
  linkcolor=blue,
  filecolor=magenta,
  urlcolor=cyan,
}

%%% symbols, notations, etc.
\usepackage{physics,braket,bm,commath,amssymb} % physics and math
\renewcommand{\t}{\text} % text in math mode
\newcommand{\f}[2]{\dfrac{#1}{#2}} % shorthand for fractions
\newcommand{\p}[1]{\left(#1\right)} % parenthesis
\renewcommand{\sp}[1]{\left[#1\right]} % square parenthesis
\renewcommand{\set}[1]{\left\{#1\right\}} % curly parenthesis
\renewcommand{\v}{\bm} % bold vectors
\newcommand{\uv}[1]{\hat{\v{#1}}} % unit vectors
\renewcommand{\d}{\partial} % partial d
\renewcommand{\c}{\cdot} % inner product
\newcommand{\bk}{\Braket} % shorthand for braket notation

\usepackage[inline]{enumitem} % for inline enumeration

% leave a note in the text, visible in the compiled document
\newcommand{\note}[1]{\textcolor{red}{#1}}
\usepackage{ulem} % for strikeout text

\usepackage{dsfont}
\newcommand{\1}{\mathds{1}}

\newcommand{\up}{\uparrow}
\newcommand{\dn}{\downarrow}
\newcommand{\E}{\mathcal{E}}
\renewcommand{\H}{\mathcal{H}}
\newcommand{\K}{\mathcal{K}}
\newcommand{\Z}{\mathbb{Z}}

\usepackage{accents}
\newcommand{\utilde}[1]{\underaccent{\tilde}{#1}}

\newcommand{\g}{\text{g}}
\newcommand{\e}{\text{e}}

\begin{document}

\title{Interactions in the 1-D optical lattice clock}

\author{Michael A. Perlin}

\maketitle

The interaction Hamiltonian for nuclear-spin polarized fermionic
alkali-earth(-like) atoms is
\begin{align}
  H_{\t{int}} = G \int d^3x~
  \psi_\e^\dag\p{x} \psi_\g^\dag\p{x} \psi_\g\p{x} \psi_\e\p{x},
  &&
  G \equiv \f{4\pi a_{\e\g^-}}{m_A},
\end{align}
where $\psi_\sigma$ is a fermionic field operator for atoms in
electronic state $s\in\set{\g,\e}$, $a_{eg}^-$ is a scattering length,
and $m_A$ is the mass of a single atom.  On a 1-D lattice with tight
transverse confinement, we can expand the field operators as
\begin{align}
  \psi_\sigma\p{x} = \sum_{q,n} \phi_{qn}\p{x} c_{qn\sigma},
\end{align}
for wavefunctions $\phi_{qn}$ and annihilation operators
$c_{qn\sigma}$ additionally indexed by quasi-momenta $q$ and band
indices $n$.  The interaction Hamiltonian then becomes
\begin{align}
  H_{\t{int}} = G \sum K^{pk;q\ell}_{rm;sn}
  c_{rm,\e}^\dag c_{sn,\g}^\dag c_{q\ell,\g} c_{pk,\e},
  &&
  K^{pk;q\ell}_{rm;sn} \equiv \int d^3x~
  \phi_{rm}\p{x}^* \phi_{sn}\p{x}^* \phi_{q\ell}\p{x} \phi_{pk}\p{x}.
\end{align}
By conservation of momentum, we know that $K^{pk;q\ell}_{rm;sn}=0$
unless $p+q+r+s=0$ (where we note that all quasi-momenta are defined
modulo 2 in units of the lattice wavenumber), so we can write
\begin{align}
  H_{\t{int}} = G \sum K^{pk;q\ell}_{p+r,m;q-r,n}
  c_{p+r,m,\e}^\dag c_{q-r,n,\g}^\dag c_{q\ell,\g} c_{pk,\e}
  \approx G \sum K^{k\ell}_{mn}
  c_{p+r,m,\e}^\dag c_{q-r,n,\g}^\dag c_{q\ell,\g} c_{pk,\e},
  \label{eq:H_int_full}
\end{align}
where we made the approximation that $K^{pk;q\ell}_{p+r,m;q-r,n}$ only
weakly depends on the quasi-momenta $\p{p,q,r}$.  We note that for the
cases we will be considering, this approximation breaks down at the
$\sim10\%$ level for lattices with depth $\lesssim 15 E_R$, where
$E_R$ is the lattice recoil energy.

\section{Single-band spin model}

We now restrict ourselves to considering only atoms a single band,
such that $K^{k\ell}_{mn}\to K$ and the single-particle energy (up to
a global shift) takes the form $E_q=-2J\cos\p{\pi q}$, where we measure
momenta $q$ in units of the lattice wavenumber.  If
$\abs{J}\gg\abs{GK}$, then by the secular approximation we can neglect
terms in $H_{\t{int}}$ which do not conserve the sum of
single-particle energies.  By solving the single-particle energy
conservation condition, one can find that this approximation amounts
to neglecting terms with $\set{p,q}\ne\set{p+r,q-r}$; that is, atoms
in different clock states can only interact via
\begin{enumerate*}[label=(\roman*)]
\item direct density-density terms, and
\item terms which exchange their momenta.
\end{enumerate*}
The surviving terms in \eqref{eq:H_int_full} are thus
\begin{align}
  H_{\t{int}}^{\t{single-band}}
  = G K \sum \p{c_{p,\e}^\dag c_{q,\g}^\dag c_{q,\g} c_{p,\e}
    + c_{q,\e}^\dag c_{p,\g}^\dag c_{q,\g} c_{p,\e}}
  = G K \sum \p{n_{q,\g} n_{p,\e}
    - c_{q,\e}^\dag c_{q,\g} c_{p,\g}^\dag c_{p,\e}}.
\end{align}
Defining
\begin{align}
  \1_p \equiv c_{p,\e}^\dag c_{p,\e} + c_{p,\g}^\dag c_{p,\g},
  &&
  \sigma_p^j \equiv \sum_{\alpha,\beta}
  c_{p\alpha}^\dag \sigma^j_{\alpha\beta} c_{p\beta},
\end{align}
for Pauli matrices $\sigma^j$, we can therefore write
\begin{align}
  H_{\t{int}}^{\t{single-band}} = G K \sum
  \sp{\p{\f{\1_q+\sigma_q^z}{2}} \p{\f{\1_p-\sigma_p^z}{2}}
    - \sigma_q^- \sigma_p^+}.
\end{align}
Up to a global shift in energy, this result is equivalently
\begin{align}
  H_{\t{int}}^{\t{single-band}}
  = - \f14 G K \sum \v\sigma_q\c\v\sigma_p
  = - G K \v S \c \v S,
  &&
  \v S \equiv \f12 \sum_p \v\sigma_p.
\end{align}
Note that the collective spin operators $\v S$ are extensive in the
number of particles, $N$, and that the overlap integral $K\sim L^{-1}$
for $L$ lattice sites, as
\begin{align}
  K = \int d^3x \abs{\phi_0\p{x}}^4
  = \int d^3x \abs{\f1{\sqrt{L}}\sum_j w_j\p{x}}^4
  \approx \f1{L^2} \sum_j \int d^3x \abs{w_j\p{x}}^4
  = \f1L \int d^3x \abs{w_0\p{x}}^4,
\end{align}
where $w_j\p{x}$ is a ground-band Wannier orbital localized at site
$j$, and the remaining overlap integral is independent of $L$.  The
approximation above is equivalent to neglecting inter-site
interactions and interaction-assisted hopping.  If $L\sim N$ then the
single-band interaction Hamiltonian is extensive in the number of
particles ($N$), as one should expect.

\section{Inter-band interactions}

We now consider interactions between atoms in two different bands, and
define $a_{p\sigma} \equiv c_{p,0,\sigma}$ (for atoms in band 0) and
$b_{p\sigma} \equiv c_{p,1,\sigma}$ (for atoms in band 1).  The
single-particle energies in the first two bands take the form
$E_{qn}=\bar E_n+2J_n\cos\p{\pi q}$ with
$\abs{\bar E_1-\bar E_0}\gg\abs{J_1}\gg\abs{J_0}$.  In the weakly
interacting limit, conserving single-particle energies requires
conserving particle number within each band, so the relevant
interactions are
\begin{align}
  H_{\t{int}}^{\t{inter-band}}
  = G K_{0,1} \sum
  \p{a_{p+r,\sigma}^\dag b_{q-r,\bar\sigma}^\dag
    + b_{p+r,\sigma}^\dag a_{q-r,\bar\sigma}^\dag}
  b_{q\bar\sigma} a_{p\sigma}
\end{align}
for $K_{0,1}\equiv K^{0,1}_{0,1}=-K^{0,0}_{1,1}>0$.  In words: atoms
in different bands may either remain in their respective clock states,
or swap a clock excitation.

The band curvatures from nonzero $J_0,J_1$ further restrict allowed
momentum changes (i.e. values of $r$) from inter-band interactions by
imposing additional constraints to conserve net single-particle
energy.  In the limit $\abs{J_1}\gg\abs{J_0}$, we find that the only
terms which conserve net single-particle energy are proportional to
$b_{q\tau}^\dag b_{q\sigma}$ and $b_{-q,\tau}^\dag b_{q\sigma}$, which
gives us
\begin{multline}
  H_{\t{int}}^{\t{inter-band}}
  = G K_{0,1} \sum
  \p{a_{p\sigma}^\dag b_{q\bar\sigma}^\dag
    + b_{q\sigma}^\dag a_{p\bar\sigma}^\dag
    + a_{p+2q,\sigma}^\dag b_{-q,\bar\sigma}^\dag
    + b_{-q,\sigma}^\dag a_{p+2q,\bar\sigma}^\dag}
  b_{q\bar\sigma} a_{p\sigma} \\
  = G K_{0,1} \sum
  \p{b_{q\bar\sigma}^\dag b_{q\bar\sigma} a_{p\sigma}^\dag a_{p\sigma}
    - b_{q\sigma}^\dag b_{q\bar\sigma} a_{p\bar\sigma}^\dag a_{p\sigma}
    + b_{-q,\bar\sigma}^\dag b_{q\bar\sigma} a_{p+2q,\sigma}^\dag a_{p\sigma}
    - b_{-q,\sigma}^\dag b_{q\bar\sigma} a_{p+2q,\bar\sigma}^\dag a_{p\sigma}}.
\end{multline}
The first two terms in this sum
($\sim b_{q\bar\sigma}^\dag b_{q\bar\sigma}$ and
$b_{q\sigma}^\dag b_{q\bar\sigma}$) correspond to density-density
terms and an exchange of clock state between atoms in different bands.
Unlike the single-band case, momentum-exchanging processes are now
disallowed due to unequal band curvatures: the fact that $J_0\ne J_1$
implies $E_{p,0}+E_{q,1}\ne E_{q,0}+E_{p,1}$.  The latter terms in the
sum ($\sim b_{-q,\bar\sigma}^\dag b_{q\bar\sigma}$ and
$b_{-q,\sigma}^\dag b_{q\bar\sigma}$) appear because
$E_{q,1}=E_{-q,1}$, i.e. the excited-band atom pays no energetic
penalty for the momentum change $q\to-q$.  To conserve net momentum in
such a process, the ground-band atom must simultaneously change
momentum as $p\to p+2q$.  At face value, this process does not
conserve energy: $E_{p,0}\ne E_{p+2q,0}$, and in turn
$E_{p,0}+E_{q,1}\ne E_{p+2q,0}+E_{-q,1}$.  For this reason, such a
process was not allowed in the single-band case.  In the two-band case
with $\abs{J_1}\gg\abs{J_0}$, however, small changes in the momentum
of the excited-band atom result in large changes in its
single-particle energy (i.e. large on the scale of conserved energy
violation, which is $\sim J_0$).  A small momentum shift
$-q\to-q+\delta q$ (together with a corresponding shift
$p+2q\to p+2q-\delta q$) is therefore sufficient to restore the
initial single-particle energy of the two atoms.

We can simplify the first set of terms of the inter-band Hamiltonian
similarly to the way that we did in the single-band case by defining
collective (pseudo-)spin operators $\v S_n$ for band $n$.  Up to a
global shift in energy, we can then take
\begin{align}
  \sum \p{b_{q\bar\sigma}^\dag b_{q\bar\sigma} a_{p\sigma}^\dag a_{p\sigma}
    - b_{q\sigma}^\dag b_{q\bar\sigma} a_{p\bar\sigma}^\dag a_{p\sigma}}
  \to -2\v S_0\c\v S_1,
\end{align}
which implies
\begin{align}
  H_{\t{int}}^{\t{inter-band}}
  = -2 G K_{0,1} \v S_0\c\v S_1
  + G K_{0,1} \sum
  \p{b_{-q,\bar\sigma}^\dag b_{q\bar\sigma} a_{p+2q,\sigma}^\dag a_{p\sigma}
    - b_{-q,\sigma}^\dag b_{q\bar\sigma} a_{p+2q,\bar\sigma}^\dag a_{p\sigma}}.
\end{align}

\section{Two-atom band-hopping}

If we use driving schemes to make $\bar E_1=\bar E_0$ (discussed in a
separate set of notes), we must also consider interaction terms which
do not conserve particle number within each band.  Such terms take the
form
\begin{align}
  H_{\t{int}}^{\t{band-hopping}} = - G K_{0,1}
  \sum \p{b_{p+r,\e}^\dag b_{q-r,\g}^\dag a_{q,\g} a_{p,\e} + \t{h.c.}}.
  \label{eq:H_int_band_hopping}
\end{align}
Energy conservation in the limit $\abs{J_1}\gg\abs{J_0}$ now forces
$r=\p{q-p}/2\pm1/2$, such that the single-particle energies of
excited-band states addressed by $H_{\t{int}}^{\t{band-hopping}}$ are
\begin{multline}
  2J_1\sp{\cos\p{\pi\sp{p+r}} + \cos\p{\pi\sp{q-r}}}
  = 2J_1\sp{\cos\p{\pi~\f{p+q}{2}+\f{\pi}{2}}
    + \cos\p{\pi~\f{p+q}{2}-\f{\pi}{2}}} \\
  = 2J_1\sp{-\sin\p{\pi~\f{p+q}{2}} + \sin\p{\pi~\f{p+q}{2}}}
  = 0.
\end{multline}
Similarly to the situation we encountered in the case of inter-band
interactions, at face value a constant single-particle energy of all
excited-band states generally implies that energy is not conserved by
the band-hopping processes in \eqref{eq:H_int_band_hopping}.  Due to
the strong dependence of excited-band single-particle energies
$E_{q,1}$ on quasi-momenta $q$, however, energy conservation can be
restored by a small shift $r\to r+\delta r$.  Keeping only terms in
\eqref{eq:H_int_band_hopping} with $r=\p{q-p}/2\pm1/2$, we thus find
\begin{align}
  H_{\t{int}}^{\t{band-hopping}}
  = - \sqrt{2} G K_{0,1} \sum
  \p{d_{\p{p+q}/2}^\dag a_{q,\g} a_{p,\e} + \t{h.c.}},
\end{align}
where
\begin{align}
  d_s^\dag \equiv \f1{\sqrt2}
  \p{b_{s+1/2,\e}^\dag b_{s-1/2,\g}^\dag
    + b_{s-1/2,\e}^\dag b_{s+1/2,\g}^\dag}.
\end{align}
For $s\ne t$, these two-body excited-band operators satisfy the
commutation relations of hard-core bosons, i.e.
\begin{align}
  \sp{d_s, d_t} = \sp{d_s^\dag, d_t^\dag} = \sp{d_s, d_t^\dag} = 0,
\end{align}
\begin{align}
  \sp{d_s, d_s^\dag}
  = 1 - \f12 \p{\tilde n_{s+1/2,\e}^\dag + \tilde n_{s-1/2,\e}^\dag
    + \tilde n_{s+1/2,\g}^\dag + \tilde n_{s-1/2,\g}^\dag},
  &&
  \tilde n_{p\sigma} \equiv b_{p\sigma}^\dag b_{p\sigma}.
\end{align}

\section{Net interaction Hamiltonian}

Putting everything together and letting $G_X\equiv G K_X$, our total
interaction Hamiltonian is
\begin{align}
  H_{\t{int}}
  &= - G_0\v S_0\c\v S_0 - G_1\v S_1\c\v S_1 - 2 G_{0,1}\v S_0\c\v S_1
  \tag*{} \\ &\quad + G_{0,1} \sum
  \p{b_{-q,\bar\sigma}^\dag b_{q\bar\sigma} a_{p+2q,\sigma}^\dag a_{p\sigma}
    - b_{-q,\sigma}^\dag b_{q\bar\sigma} a_{p+2q,\bar\sigma}^\dag a_{p\sigma}}
  \tag*{} \\ &\quad - \sqrt2 G_{0,1} \sum
  \p{d_{\p{p+q}/2}^\dag a_{q,\g} a_{p,\e} + \t{h.c.}},
\end{align}
where
\begin{align}
  d_s^\dag \equiv \f1{\sqrt2}
  \p{b_{s+1/2,\e}^\dag b_{s-1/2,\g}^\dag
    + b_{s-1/2,\e}^\dag b_{s+1/2,\g}^\dag}.
\end{align}
Notes:
\begin{enumerate}[label=(\roman*)]
\item The last set of terms ($\sim d_s^\dag$ and $d_s$) appear only if
  we use driving schemes to match the mean band energies, thereby
  effective setting $\bar E_1=\bar E_0$; these terms can be neglected
  in the absence of such driving.
\item The number of nonvanishing terms in $\v S_n$ is proportional to
  the occupation of band $n$.  If there is a large population of atoms
  in the ground band and a small population of atoms in the excited
  band, then we can safely neglect the $\sim\v S_1\c\v S_1$ term,
  which is doubly small compared to e.g. $\v S_0\c\v S_1$.
\item Similarly, if the ground band is nearly filled with atoms in one
  clock state and the excited band only sparsely populated with atoms
  in the opposite clock state, then we can neglect the terms
  $\sim b_{-q,\bar\sigma}^\dag b_{q\bar\sigma} a_{p+2q,\sigma}^\dag
  a_{p\sigma}$, as $b_{-q,\bar\sigma}^\dag b_{q\bar\sigma}$ manifestly
  addresses a small number of atoms while
  $a_{p+2q,\sigma}^\dag a_{p\sigma}$ is only nonzero for quasi-momenta
  $p$ with $p+2q$ unoccupied.
\item If there is a nonzero population of atoms in the excited band,
  the terms
  $\sim b_{-q,\sigma}^\dag b_{q\bar\sigma} a_{p+2q,\bar\sigma}^\dag
  a_{p\sigma}$ will generally cause delocalization of clock
  excitations across all quasi-momenta (on a time scale
  $\sim1/G_{0,1}$ which is extensive with $N$).
\end{enumerate}

\end{document}

### 2018-07-09 link_model/link_model_notes.tex
added quantum link model notes

\documentclass[aps,nofootinbib,notitlepage,11pt]{revtex4-1}

% linking references
\usepackage{hyperref}
\hypersetup{
  breaklinks=true,
  colorlinks=true,
  linkcolor=blue,
  filecolor=magenta,
  urlcolor=cyan,
}

%%% symbols, notations, etc.
\usepackage{physics,braket,bm,amssymb} % physics and math
\renewcommand{\t}{\text} % text in math mode
\newcommand{\f}[2]{\dfrac{#1}{#2}} % shorthand for fractions
\newcommand{\p}[1]{\left(#1\right)} % parenthesis
\renewcommand{\sp}[1]{\left[#1\right]} % square parenthesis
\renewcommand{\set}[1]{\left\{#1\right\}} % curly parenthesis
\renewcommand{\v}{\bm} % bold vectors
\newcommand{\uv}[1]{\hat{\v{#1}}} % unit vectors
\newcommand{\av}{\vec} % arrow vectors
\newcommand{\del}{\nabla} % del operator
\renewcommand{\d}{\partial} % partial d
\renewcommand{\c}{\cdot} % inner product
\newcommand{\w}{\wedge} % wedge product
\newcommand{\bk}{\Braket} % shorthand for braket notation

\newcommand{\up}{\uparrow}
\newcommand{\dn}{\downarrow}
\newcommand{\g}{\text{g}}
\newcommand{\e}{\text{e}}
\renewcommand{\L}{\text{L}}
\newcommand{\C}{\text{C}}
\newcommand{\R}{\text{R}}
\newcommand{\B}{\text{B}}
\renewcommand{\P}{\text{P}}
\newcommand{\T}{\text{T}}

\usepackage{dsfont} % for \1
\newcommand{\1}{\mathds{1}}

\usepackage{mathtools} % for coloneqq
\usepackage{tikz} % for energy level diagram
\tikzset{
  baseline = (current bounding box.center),
}
\usetikzlibrary{decorations.pathmorphing}

\usepackage[inline]{enumitem} % in-line lists

% leave a note in the text, visible in the compiled document
\newcommand{\note}[1]{\textcolor{red}{#1}}

\begin{document}

\title{Basic operations for a quantum link model with ultracold
  $^{87}$Sr on a lattice}

\author{Michael A. Perlin}

\maketitle

We consider loading ultracold $^{87}$Sr atoms into an isolated
three-site cell of a state-dependent 2-D optical superlattice, with
real on-sites wavefunctions $\phi_\L$ (left) and $\phi_\R$ (right) for
atoms in the ground electronic state ($^1S_0$), and $\phi_\C$ (center)
for atoms in the excited electronic state ($^3P_0$).  The
single-particle Hamiltonian is then
\begin{align}
  H_0 = \sum_{s,\mu} E_s c_{s\mu}^\dag c_{s\mu},
\end{align}
for sites $r,s\in\set{\L,\C,\R}$, nuclear spins
$\mu\in\set{-I,\cdots,I}$, and fermionic annihilation operators
$c_{s\mu}$.  The energy mismatches $\Delta_{pq}\equiv E_p-E_q$ between
sites $p\ne q$ will generally set the largest energy scale throughout
this work.  If there are multiple atoms loaded within the same cell,
these atoms interact through the two-body Hamiltonian
\begin{align}
  H_{\t{int}} = \f12 \sum_{\substack{p,q,r,s\\\mu,\nu}}
  V^{pq}_{rs} c_{r\mu}^\dag c_{s\nu}^\dag c_{q\nu} c_{p\mu}
  \label{eq:H_int_full}
\end{align}
for interaction strengths $V^{pq}_{rs}$ which depend only on the sites
and their associated electronic states; these interaction strengths do
not depend on the nuclear spin states involved.  Defining
\begin{align}
  K^{pq}_{rs} \equiv \int d^3x~ \phi_r^* \phi_s^* \phi_q \phi_p,
  &&
  G_{\sigma=\g,\e} \equiv \f{4\pi}{m_A} a_{\sigma\sigma},
  &&
  G_\pm \equiv \f{2\pi}{m_A} \p{a_{\e\g+} \pm a_{\e\g-}},
\end{align}
\begin{align}
  G^{\alpha\beta}_{\gamma\delta} \equiv \left\{
    \begin{array}{ll}
      G_\alpha & ~ \alpha = \beta = \gamma = \delta \\
      G_+ & ~ \alpha \ne \beta ~ \t{and} ~
            (\alpha,\beta) = (\gamma,\delta) \\
      G_- & ~ \alpha \ne \beta ~ \t{and} ~
            (\alpha,\beta) = (\delta,\gamma) \\
      0 & ~ \t{otherwise}
    \end{array}\right.,
\end{align}
for an atomic mass $m_A$ and scattering lengths $a_X$, the interaction
strengths in \eqref{eq:H_int_full} are
\begin{align}
  V^{pq}_{rs} = K^{pq}_{rs}
  G^{\varepsilon\p{p}\varepsilon\p{q}}_{\varepsilon\p{r}\varepsilon\p{s}},
\end{align}
where $\varepsilon\p{p}\in\set{\g,\e}$ selects the electronic state
associated with site $p$.  As a first approximation, we assume that
the two-body overlap integral $K^{pq}_{rs}$ is negligible unless
$\set{p,q}=\set{r,s}$, i.e. we neglect interaction-assisted tunneling
in $H_{\t{int}}$.  In the presence of only $H_0$ and $H_{\t{int}}$,
this approximation can also be justified by the secular approximation
due to mismatched single-particle on-site energies $E_s$.  The
interaction Hamiltonian then simplifies to
\begin{align}
  H_{\t{int}}
  = \f12 \sum_{p,\mu,\nu} V_p c_{p\mu}^\dag c_{p\nu}^\dag c_{p\nu} c_{p\mu}
  + \f12 \sum_{\substack{p\ne q\\\mu,\nu}} V_{pq}
  \p{c_{p\mu}^\dag c_{q\nu}^\dag c_{q\nu} c_{p\mu}
    + c_{q\mu}^\dag c_{p\nu}^\dag c_{q\nu} c_{p\mu}},
  \label{eq:H_int_two}
\end{align}
where
\begin{align}
  V_p \equiv V^{pp}_{pp},
  &&
  V_{pq} \equiv V^{pq}_{pq} = V^{pq}_{qp},
  &&
  n_{p\mu} \equiv c_{p\mu}^\dag c_{p\mu}.
\end{align}
As made clear by the form in \eqref{eq:H_int_two}, the interaction
Hamiltonian $H_{\t{int}}$ generally induces nuclear spin exchange
between atoms in different orbitals.  In order to suppress nuclear
spin exchange, we further restrict ourselves to a regime in which the
two-body overlap integral $K^{pq}_{rs}$ is negligible unless all atoms
are on the same site (i.e. unless $p=q=r=s$), such that
\begin{align}
  H_{\t{int}}
  = \f12 \sum_{p,\mu,\nu} V_p c_{p\mu}^\dag c_{p\nu}^\dag c_{p\nu} c_{p\mu}
  = \sum_{\substack{p\\\mu<\nu}} V_p n_{p\mu} n_{p\nu}.
  \label{eq:H_int}
\end{align}
Throughout the remainder of this work, to simplify matters we
generally assume that atoms initially in the electronic ground state
on site $\L$ or $\R$ lie in the nuclear spin manifold spanned by
$\set{\ket{-I},\ket{-I+1}}$, while atoms initially in an excited
electronic state on site $\C$ lie in the nuclear spin manifold spanned
by $\set{\ket{I},\ket{I-1}}$.  We also assume that each site is
initially populated with at most one atom.

\section{Global control gates}
\label{sec:global_control}

We wish to build a toolkit of operations to manipulate atoms in our
three-site cell, with the long term goal of combining many such cells
to simulate a quantum link model.  Before we discuss any multi-body
gates, we first need to construct some basic, single-body global
control operations.  These basic operations together with the two-body
interaction Hamiltonian $H_{\t{int}}$ in \eqref{eq:H_int} will be our
building blocks for realizing more complicated gates.

The first simple operation we consider is the application of a static
magnetic field with strength $B$ along the nuclear spin quantization
axis (which we will generally keep fixed), realizing the Hamiltonian
\begin{align}
  H_\B = -\gamma_{\text{n}} B \sum_{s,\mu} \mu n_{s\mu},
\end{align}
where $\gamma_{\text{n}}$ is the nuclear gyromagnetic ratio.  Applying
this Hamiltonian for a time $t=\theta/\gamma_nB$ generates the phase
gate
\begin{align}
  U_\B\p{\theta} = \sum_{s,\mu} e^{i \theta \mu} n_{s\mu}.
\end{align}
As the nuclear spins $\mu$ are half-integers, the operation
$U_\B\p{\theta}$ is periodic in $\theta$ with period $4\pi$.

We can also shine an auxiliary laser with angular frequency $\omega$
to apply the Hamiltonian
\begin{align}
  H_\Omega^{(\sigma)}
  = - \sum_\mu e^{i\omega t}
  \p{\Omega_{\C,\L,\mu}^{(\sigma)} c_{\L,\mu}^\dag
    + \Omega_{\C,\R,\mu}^{(\sigma)} c_{\R,\mu}^\dag} c_{\C,\mu+\sigma}
  + \t{h.c.},
\end{align}
where $\sigma\in\set{-1,0,1}$ is determined by the clock laser
polarization; $\L\leftrightarrow\R$ coupling is forbidden because
$H_\Omega^{(\sigma)}$ is strictly off-diagonal in electronic state;
and $\t{h.c.}$ denotes the Hermitian conjugate, i.e.
$X+\t{h.c.}\equiv X+X^\dag$.  By evaluating the Clebsch-Gordan
coefficients $\bk{I,\mu+\sigma;1,-\sigma|I,\mu}$, we can extract the
$\mu,\sigma,p,q$ dependence of $\Omega_{pq\mu}^{(\sigma)}$ to find
\begin{align}
  \Omega_{pq\mu}^{(0)} = \Omega_{pq}^{(0)} \mu,
  &&
  \Omega_{pq\mu}^{(\pm1)} = \pm \Omega_{pq}^{(1)}
  \sqrt{\p{I\mp\mu}\p{I\pm\mu+1}},
  &&
  \Omega_{pq}^{(\sigma)} \equiv \Omega_\sigma \int d^3x~ \phi_q^* \phi_p,
\end{align}
for experimentally tunable intensities $\Omega_0$ and $\Omega_1$ which
are independent of nuclear spin and lattice site.  In this section, we
will generally assume that these intensities are real and large
compared to all interaction energies,
i.e. $\Omega_0,\Omega_1\gg\max_p\abs{V_p}$.  We also consider the case
of ${}^{87}$Sr with net nuclear spin $I=9/2$ in order to fix the
relative magnitudes of couplings $\Omega_{pq\mu}^{(\sigma)}$ for
different nuclear spins $\mu$.  The Hamiltonians $H_\Omega^{(\sigma)}$
can be thought of diagrammatically as
\begin{align}
  H_\Omega^{(0)} \sim
  \begin{tikzpicture}[
    scale=0.5,
    trans/.style={thick,<->,shorten >=2pt,shorten <=2pt,>=stealth}
    ]
    \draw (0,0) -- (1,0);
    \draw (2,0) -- (3,0);
    \draw[dotted] (3.5,0) -- (4.5,0);
    \draw (5,0) -- (6,0);
    \draw (7,0) -- (8,0);
    \draw (7,0) -- (8,0) node[right] {$\C$};
    \draw (0,-2) -- (1,-2);
    \draw (2,-2) -- (3,-2);
    \draw[dotted] (3.5,-2) -- (4.5,-2);
    \draw (5,-2) -- (6,-2);
    \draw (7,-2) -- (8,-2);
    \draw (7,-2) -- (8,-2) node[right] {$s\in\set{\L,\R}$};
    \draw[trans] (0.5,0) -- (0.5,-2) node[midway,left] {$-9$};
    \draw[trans] (2.5,0) -- (2.5,-2) node[midway,left] {$-7$};
    \draw[trans] (5.5,0) -- (5.5,-2) node[midway,left] {$7$};
    \draw[trans] (7.5,0) -- (7.5,-2) node[midway,left] {$9$};
  \end{tikzpicture},
\end{align}
where the horizontal lines represent nuclear spin levels from $-9/2$
to $9/2$ (i.e. from left to right) within each orbital state $\C$ or
$s\in\set{\L,\R}$, and the double-headed arrows represent a coupling
between two states.  The numbers next to the arrows represents the
relative magnitudes of the couplings.  Similarly,
\begin{align}
  H_\Omega^{(1)} \sim
  \begin{tikzpicture}[
    scale=0.5,
    trans/.style={thick,<->,shorten >=2pt,shorten <=2pt,>=stealth}
    ]
    \draw (0,0) -- (1,0);
    \draw (2,0) -- (3,0);
    \draw (4,0) -- (5,0);
    \draw[dotted] (5.5,0) -- (6.5,0);
    \draw (7,0) -- (8,0);
    \draw (9,0) -- (10,0);
    \draw (11,0) -- (12,0);
    \draw (11,0) -- (12,0) node[right] {$\C$};
    \draw (0,-2) -- (1,-2);
    \draw (2,-2) -- (3,-2);
    \draw (4,-2) -- (5,-2);
    \draw[dotted] (5.5,-2) -- (6.5,-2);
    \draw (7,-2) -- (8,-2);
    \draw (9,-2) -- (10,-2);
    \draw (11,-2) -- (12,-2);
    \draw (11,-2) -- (12,-2) node[right] {$s\in\set{\L,\R}$};
    \draw[trans] (0.5,-2) -- (2.5,0) node[midway,left] {$3$};
    \draw[trans] (2.5,-2) -- (4.5,0) node[midway,left] {$4$};
    \draw[trans] (7.5,-2) -- (9.5,0) node[midway,left] {$4$};
    \draw[trans] (9.5,-2) -- (11.5,0) node[midway,left] {$3$};
  \end{tikzpicture},
  \label{eq:H_O_+}
\end{align}
and
\begin{align}
  H_\Omega^{(-1)} \sim
  \begin{tikzpicture}[
    scale=0.5,
    trans/.style={thick,<->,shorten >=2pt,shorten <=2pt,>=stealth}
    ]
    \draw (0,0) -- (1,0);
    \draw (2,0) -- (3,0);
    \draw (4,0) -- (5,0);
    \draw[dotted] (5.5,0) -- (6.5,0);
    \draw (7,0) -- (8,0);
    \draw (9,0) -- (10,0);
    \draw (11,0) -- (12,0);
    \draw (11,0) -- (12,0) node[right] {$\C$};
    \draw (0,-2) -- (1,-2);
    \draw (2,-2) -- (3,-2);
    \draw (4,-2) -- (5,-2);
    \draw[dotted] (5.5,-2) -- (6.5,-2);
    \draw (7,-2) -- (8,-2);
    \draw (9,-2) -- (10,-2);
    \draw (11,-2) -- (12,-2);
    \draw (11,-2) -- (12,-2) node[right] {$s\in\set{\L,\R}$};
    \draw[trans] (0.5,0) -- (2.5,-2) node[midway,right] {$3$};
    \draw[trans] (2.5,0) -- (4.5,-2) node[midway,right] {$4$};
    \draw[trans] (7.5,0) -- (9.5,-2) node[midway,right] {$4$};
    \draw[trans] (9.5,0) -- (11.5,-2) node[midway,right] {$3$};
  \end{tikzpicture},
  \label{eq:H_O_-}
\end{align}
where we neglect couplings for nuclear spins $\mu$ outside of
$\set{-I,-I+1,I-1,I}$.

By turning on $H_\Omega^{(0)}$ for a time $t=\pi/\Omega^{(0)}_{\C,s}$
with an angular frequency $\omega$ resonant on the energy difference
$\Delta_{\C,s}=E_C-E_s$ for $s\in\set{\L,\R}$, we can apply the
$\pi$-pulse gate
\begin{align}
  U_{\pi,s}
  \equiv \sum_\mu e^{i\pi\mu} c_{s\mu}^\dag c_{\C,\mu} + \t{h.c}
  \coloneqq i \times
  \begin{tikzpicture}[
    scale=0.5,
    trans/.style={thick,<->,shorten >=2pt,shorten <=2pt,>=stealth}
    ]
    \draw (0,0) -- (1,0);
    \draw (2,0) -- (3,0);
    \draw[dotted] (3.5,0) -- (4.5,0);
    \draw (5,0) -- (6,0);
    \draw (7,0) -- (8,0);
    \draw (7,0) -- (8,0) node[right] {$\C$};
    \draw (0,-2) -- (1,-2);
    \draw (2,-2) -- (3,-2);
    \draw[dotted] (3.5,-2) -- (4.5,-2);
    \draw (5,-2) -- (6,-2);
    \draw (7,-2) -- (8,-2);
    \draw (7,-2) -- (8,-2) node[right] {$s$};
    \draw[trans] (0.5,0) -- (0.5,-2) node[midway,left] {$-1$};
    \draw[trans] (2.5,0) -- (2.5,-2) node[midway,left] {$1$};
    \draw[trans] (5.5,0) -- (5.5,-2) node[midway,left] {$-1$};
    \draw[trans] (7.5,0) -- (7.5,-2) node[midway,left] {$1$};
  \end{tikzpicture},
  \label{eq:U_pi}
\end{align}
which swaps atoms between orbitals $\C$ and $s$ and applies a
spin-dependent sign.

The relative magnitudes of the couplings $\Omega_{pq\mu}^{(\pm1)}$ for
different nuclear spins $\mu$ similarly allows us to turn on
$H_\Omega^{(\pm1)}$ for an appropriately resonant angular frequency
$\omega$ and a time $t=\p{\pi/2}/\Omega_{\C,s}^{(1)}$, thereby
realizing the gates
\begin{align}
  U_{\pi,s}^+
  \equiv -i\p{c_{s,-I+1} c_{\C,-I+1}^\dag + c_{s,I-1}^\dag c_{\C,I}}
  + \t{h.c.}
  \coloneqq -i \times
  \begin{tikzpicture}[
    scale=0.5,
    trans/.style={thick,<->,shorten >=2pt,shorten <=2pt,>=stealth}
    ]
    \draw (0,0) -- (1,0);
    \draw (2,0) -- (3,0);
    \draw[dotted] (3.5,0) -- (4.5,0);
    \draw (5,0) -- (6,0);
    \draw (7,0) -- (8,0);
    \draw (7,0) -- (8,0) node[right] {$\C$};
    \draw (0,-2) -- (1,-2);
    \draw (2,-2) -- (3,-2);
    \draw[dotted] (3.5,-2) -- (4.5,-2);
    \draw (5,-2) -- (6,-2);
    \draw (7,-2) -- (8,-2);
    \draw (7,-2) -- (8,-2) node[right] {$s$};
    \draw[trans] (2.5,0) -- (0.5,-2);
    \draw[trans] (7.5,0) -- (5.5,-2);
  \end{tikzpicture},
  \label{eq:U_pi_+}
\end{align}
and
\begin{align}
  U_{\pi,s}^-
  \equiv i\p{c_{s,-I+1} c_{\C,-I+1}^\dag + c_{s,I}^\dag c_{\C,I-1}}
  + \t{h.c.}
  \coloneqq i \times
  \begin{tikzpicture}[
    scale=0.5,
    trans/.style={thick,<->,shorten >=2pt,shorten <=2pt,>=stealth}
    ]
    \draw (0,0) -- (1,0);
    \draw (2,0) -- (3,0);
    \draw[dotted] (3.5,0) -- (4.5,0);
    \draw (5,0) -- (6,0);
    \draw (7,0) -- (8,0);
    \draw (7,0) -- (8,0) node[right] {$\C$};
    \draw (0,-2) -- (1,-2);
    \draw (2,-2) -- (3,-2);
    \draw[dotted] (3.5,-2) -- (4.5,-2);
    \draw (5,-2) -- (6,-2);
    \draw (7,-2) -- (8,-2);
    \draw (7,-2) -- (8,-2) node[right] {$s$};
    \draw[trans] (0.5,0) -- (2.5,-2);
    \draw[trans] (5.5,0) -- (7.5,-2);
  \end{tikzpicture},
  \label{eq:U_pi_-}
\end{align}
where the remaining states addressed by $H_\Omega^{(\pm1)}$ in
\eqref{eq:H_O_+} and \eqref{eq:H_O_-} are unaffected because they
undergo Rabi cycles which bring them back to their initial states.

All of the above $\pi$-pulse gates can be turned into $\pi/2$-pulse
gates $U_{\pi/2,s}$, $U_{\pi/2,s}^\pm$ simply by applying the
appropriate Hamiltonian for half of the specified time.  Furthermore,
if we are guaranteed that only certain nuclear spin states are
occupied, we can also apply $\pi$-pulse gates by turning on
$H_\Omega^{(\pm1)}$ for a time $t=\p{\pi/6}/\Omega_{\C,s}^{(1)}$,
giving us
\begin{align}
  \tilde U_{\pi,s}^+
  \coloneqq i \times
  \begin{tikzpicture}[
    scale=0.5,
    trans/.style={thick,<->,shorten >=2pt,shorten <=2pt,>=stealth}
    ]
    \draw (0,0) -- (1,0);
    \draw (2,0) -- (3,0);
    \draw[dotted] (3.5,0) -- (5.5,0);
    \draw (6,0) -- (7,0);
    \draw (8,0) -- (9,0);
    \draw (8,0) -- (9,0) node[right] {$\C$};
    \draw (0,-2) -- (1,-2);
    \draw (2,-2) -- (3,-2);
    \draw[dotted] (3.5,-2) -- (5.5,-2);
    \draw (6,-2) -- (7,-2);
    \draw (8,-2) -- (9,-2);
    \draw (8,-2) -- (9,-2) node[right] {$s$};
    \draw[trans] (2.5,0) -- (0.5,-2);
    \draw[trans, dashed] (4.5,0) -- (2.5,-2);
    \draw[trans, dashed] (6.5,0) -- (4.5,-2);
    \draw[trans] (8.5,0) -- (6.5,-2);
  \end{tikzpicture},
  &&
  \tilde U_{\pi,s}^-
  \coloneqq -i \times
  \begin{tikzpicture}[
    scale=0.5,
    trans/.style={thick,<->,shorten >=2pt,shorten <=2pt,>=stealth}
    ]
    \draw (0,0) -- (1,0);
    \draw (2,0) -- (3,0);
    \draw[dotted] (3.5,0) -- (5.5,0);
    \draw (6,0) -- (7,0);
    \draw (8,0) -- (9,0);
    \draw (8,0) -- (9,0) node[right] {$\C$};
    \draw (0,-2) -- (1,-2);
    \draw (2,-2) -- (3,-2);
    \draw[dotted] (3.5,-2) -- (5.5,-2);
    \draw (6,-2) -- (7,-2);
    \draw (8,-2) -- (9,-2);
    \draw (8,-2) -- (9,-2) node[right] {$s$};
    \draw[trans] (2.5,-2) -- (0.5,0);
    \draw[trans, dashed] (4.5,-2) -- (2.5,0);
    \draw[trans, dashed] (6.5,-2) -- (4.5,0);
    \draw[trans] (8.5,-2) -- (6.5,0);
  \end{tikzpicture},
\end{align}
where the dashed arrows indicate incomplete Rabi cycles that mix nuclear
spin states in our target subspace with those outside it.  For
simplicity, we will work strictly with the cleaner $\pi$-pulse gates
in \eqref{eq:U_pi}, \eqref{eq:U_pi_+}, and \eqref{eq:U_pi_-}.

\section{Tunneling gates}
\label{sec:tunneling}

In order to suppress nuclear spin exchange processes, we restricted
ourselves to a regime in which two-body overlap integrals for pairs of
atoms on different sites vanish.  The on-site wavefunctions $\phi_p$
are therefore fairly localized to site $p$, with some overlap in the
support of $\phi_\C$ with $\phi_s$ for $s\in\set{\L,\R}$ to allow for
the control gates $U_{\pi,s}^\pm$.  As a consequence, we expect
vanishing mutual support of $\phi_\L$ and $\phi_\R$, which means that
we cannot induce $\L\leftrightarrow\R$ tunneling directly.
Nonetheless, we can use the $\pi$-pulse gates $U_{\pi,s}$ in
\eqref{eq:U_pi} to induce indirect $\L\leftrightarrow\R$ tunneling
through the central site $\C$ via
\begin{multline}
  U_\pi^\T
  \equiv U_{\pi,\L} U_{\pi,\R} U_{\pi,\L}
  \cong U_{\pi,\R} U_{\pi,\L} U_{\pi,\R} \\
  \cong
  \sp{\begin{tikzpicture}[
      scale=0.5,
      trans/.style={thick,<->,shorten >=2pt,shorten <=2pt,>=stealth}
      ]
      \draw (0,0) -- (1,0);
      \draw (2,0) -- (3,0);
      \draw[dotted] (3.5,0) -- (4.5,0);
      \draw (5,0) -- (6,0);
      \draw (7,0) -- (8,0);
      \draw (7,0) -- (8,0) node[right] {$\L$};
      \draw (0,-2) -- (1,-2);
      \draw (2,-2) -- (3,-2);
      \draw[dotted] (3.5,-2) -- (4.5,-2);
      \draw (5,-2) -- (6,-2);
      \draw (7,-2) -- (8,-2);
      \draw (7,-2) -- (8,-2) node[right] {$\C$};
      \draw (0,-4) -- (1,-4);
      \draw (2,-4) -- (3,-4);
      \draw[dotted] (3.5,-4) -- (4.5,-4);
      \draw (5,-4) -- (6,-4);
      \draw (7,-4) -- (8,-4);
      \draw (7,-4) -- (8,-4) node[right] {$\R$};
      \draw[trans] (0.5,0) -- (0.5,-2);
      \draw[trans] (2.5,0) -- (2.5,-2);
      \draw[trans] (5.5,0) -- (5.5,-2);
      \draw[trans] (7.5,0) -- (7.5,-2);
    \end{tikzpicture}}
  \sp{\begin{tikzpicture}[
      scale=0.5,
      trans/.style={thick,<->,shorten >=2pt,shorten <=2pt,>=stealth}
      ]
      \draw (0,0) -- (1,0);
      \draw (2,0) -- (3,0);
      \draw[dotted] (3.5,0) -- (4.5,0);
      \draw (5,0) -- (6,0);
      \draw (7,0) -- (8,0);
      \draw (7,0) -- (8,0) node[right] {$\L$};
      \draw (0,-2) -- (1,-2);
      \draw (2,-2) -- (3,-2);
      \draw[dotted] (3.5,-2) -- (4.5,-2);
      \draw (5,-2) -- (6,-2);
      \draw (7,-2) -- (8,-2);
      \draw (7,-2) -- (8,-2) node[right] {$\C$};
      \draw (0,-4) -- (1,-4);
      \draw (2,-4) -- (3,-4);
      \draw[dotted] (3.5,-4) -- (4.5,-4);
      \draw (5,-4) -- (6,-4);
      \draw (7,-4) -- (8,-4);
      \draw (7,-4) -- (8,-4) node[right] {$\R$};
      \draw[trans] (0.5,-2) -- (0.5,-4);
      \draw[trans] (2.5,-2) -- (2.5,-4);
      \draw[trans] (5.5,-2) -- (5.5,-4);
      \draw[trans] (7.5,-2) -- (7.5,-4);
    \end{tikzpicture}}
  \sp{\begin{tikzpicture}[
      scale=0.5,
      trans/.style={thick,<->,shorten >=2pt,shorten <=2pt,>=stealth}
      ]
      \draw (0,0) -- (1,0);
      \draw (2,0) -- (3,0);
      \draw[dotted] (3.5,0) -- (4.5,0);
      \draw (5,0) -- (6,0);
      \draw (7,0) -- (8,0);
      \draw (7,0) -- (8,0) node[right] {$\L$};
      \draw (0,-2) -- (1,-2);
      \draw (2,-2) -- (3,-2);
      \draw[dotted] (3.5,-2) -- (4.5,-2);
      \draw (5,-2) -- (6,-2);
      \draw (7,-2) -- (8,-2);
      \draw (7,-2) -- (8,-2) node[right] {$\C$};
      \draw (0,-4) -- (1,-4);
      \draw (2,-4) -- (3,-4);
      \draw[dotted] (3.5,-4) -- (4.5,-4);
      \draw (5,-4) -- (6,-4);
      \draw (7,-4) -- (8,-4);
      \draw (7,-4) -- (8,-4) node[right] {$\R$};
      \draw[trans] (0.5,0) -- (0.5,-2);
      \draw[trans] (2.5,0) -- (2.5,-2);
      \draw[trans] (5.5,0) -- (5.5,-2);
      \draw[trans] (7.5,0) -- (7.5,-2);
    \end{tikzpicture}} \\
  \cong
  \begin{tikzpicture}[
    scale=0.5,
    trans/.style={thick,<->,shorten >=2pt,shorten <=2pt,>=stealth}
    ]
    \draw (0,0) -- (1,0);
    \draw (2,0) -- (3,0);
    \draw[dotted] (3.5,0) -- (4.5,0);
    \draw (5,0) -- (6,0);
    \draw (7,0) -- (8,0);
    \draw (7,0) -- (8,0) node[right] {$\L$};
    \draw (0,-2) -- (1,-2);
    \draw (2,-2) -- (3,-2);
    \draw[dotted] (3.5,-2) -- (4.5,-2);
    \draw (5,-2) -- (6,-2);
    \draw (7,-2) -- (8,-2);
    \draw (7,-2) -- (8,-2) node[right] {$\R$};
    \draw[trans] (0.5,0) -- (0.5,-2);
    \draw[trans] (2.5,0) -- (2.5,-2);
    \draw[trans] (5.5,0) -- (5.5,-2);
    \draw[trans] (7.5,0) -- (7.5,-2);
  \end{tikzpicture},
  \label{eq:U_pi_T}
\end{multline}
where $\cong$ denotes equality up to an overall phase.  Crucially,
$U_\pi^\T$ swaps atoms between sites $\L$ and $\R$ while leaving atoms
on site $\C$ unaffected.  We can also construct $\pi/2$-pulse gates
for $\L\leftrightarrow\R$ tunneling via
\begin{align}
  U_{\pi/2}^\T
  \equiv U_{\pi,\L} U_{\pi/2,\R} U_{\pi,\L}
  \cong U_{\pi,\R} U_{\pi/2,\L} U_{\pi,\R},
\end{align}
which is identical to the composite gate $U_\pi^\T$ in
\eqref{eq:U_pi_T}, but with a replacement of the second $\pi$-pulse by
a $\pi/2$ pulse.  Finally, we can make use of the
nuclear-spin-selective properties of $U_{\pi,s}^\pm$ to construct
nuclear-spin-selective $\L\leftrightarrow\R$ tunneling gates, simply
by replacing instances of $U_{\pi,s}$ above by $U_{\pi,s}^\pm$.

\section{Controlled-phase gates}
\label{sec:controlled_phase}

We now construct a collisional controlled-phase gate on an atom in
site $\C$, conditional on the presence of an atom in site $s$.  We
first prepare an atom on site $\C$ in the nuclear spin manifold
spanned by $\set{\ket{I},\ket{I-1}}$, and consider the possibility of
an atom on site $s$ with nuclear spin state $\ket{-I}$, preparing the
state
\begin{align}
  \ket{\psi_{sm}}
  = \ket{s,-I;m} \otimes \p{\alpha\ket{\C,I} + \beta\ket{\C,I-1}}
  \sim
  \begin{tikzpicture}[
    scale=0.5,
    sim/.style={decorate, decoration=snake},
    ]
    \draw (0,0) -- (1,0);
    \draw (2,0) -- (3,0);
    \draw[dotted] (3.5,0) -- (4.5,0);
    \draw (5,0) -- (6,0) node[midway, above] {$\beta$};
    \draw (7,0) -- (8,0) node[midway, above] {$\alpha$};
    \draw (7,0) -- (8,0) node[right] {$\C$};
    \draw (0,-2) -- (1,-2) node[midway, above] {$m$};
    \draw (2,-2) -- (3,-2);
    \draw[dotted] (3.5,-2) -- (4.5,-2);
    \draw (5,-2) -- (6,-2);
    \draw (7,-2) -- (8,-2);
    \draw (7,-2) -- (8,-2) node[right] {$s$};
    \draw[sim] (6,0.5) -- (7,0.5);
  \end{tikzpicture}
  \label{eq:psi_sm}
\end{align}
where $m=0$ ($m=1$) marks the absence (presence) of an atom on site
$s$ with spin $-I$.  The above diagram is merely provided as a sketch
of the state; we will make no attempt at making this sketch a formal
representation.

Applying the gate $U_{\pi,s}^-$ in \eqref{eq:U_pi_-} to the state
$\ket{\psi_{sm}}$ in \eqref{eq:psi_sm} gives us
\begin{align}
  U_{\pi,s}^-\ket{\psi_{sm}}
  \cong \ket{s,-I;m} \otimes \p{\alpha\ket{\C,I} + \beta\ket{s,I}}
  \sim
  \begin{tikzpicture}[
    scale=0.5,
    sim/.style={decorate, decoration=snake},
    ]
    \draw (0,0) -- (1,0);
    \draw (2,0) -- (3,0);
    \draw[dotted] (3.5,0) -- (4.5,0);
    \draw (5,0) -- (6,0);
    \draw (7,0) -- (8,0) node[midway, above] {$\alpha$};
    \draw (7,0) -- (8,0) node[right] {$\C$};
    \draw (0,-2) -- (1,-2) node[midway, above] {$m$};
    \draw (2,-2) -- (3,-2);
    \draw[dotted] (3.5,-2) -- (4.5,-2);
    \draw (5,-2) -- (6,-2);
    \draw (7,-2) -- (8,-2) node[midway, above] {$\beta$};
    \draw (7,-2) -- (8,-2) node[right] {$s$};
    \draw[sim] (7.5,0.1) -- (7.5,-0.9);
  \end{tikzpicture}.
\end{align}
Holding for a time $t$ (i.e. evolving under $H_0+H_{\t{int}}$) now
makes $\beta$ pick up a phase relative to $\alpha$ as
$\beta\to\tilde\beta_{sm}\p{t}\equiv\beta e^{it\Delta_{\C,s}-itV_sm}$,
where $\Delta_{\C,s}=E_\C-E_s$ is the single-particle energy
difference between the orbitals $\C$ and $s$, and $V_s$ is the
interaction energy of two atoms in orbital $s$.  Subsequently applying
$U_{\pi,s}^-$ then gives us the state
\begin{align}
  U_{\pi,s}^- U_{\t{hold}}\p{t} U_{\pi,s}^- \ket{\psi_{sm}}
  \cong \ket{s,-I;m} \otimes
  \p{\alpha\ket{\C,I} + \tilde\beta_{sm}\p{t}\ket{\C,I-1}}
  \sim
  \begin{tikzpicture}[
    scale=0.5,
    sim/.style={decorate, decoration=snake},
    ]
    \draw (0,0) -- (1,0);
    \draw (2,0) -- (3,0);
    \draw[dotted] (3.5,0) -- (4.5,0);
    \draw (5,0) -- (6,0) node[near start, above]
    {$\tilde\beta_{sm}\p{t}$};
    \draw (7,0) -- (8,0) node[midway, above] {$\alpha$};
    \draw (7,0) -- (8,0) node[right] {$\C$};
    \draw (0,-2) -- (1,-2) node[midway, above] {$m$};
    \draw (2,-2) -- (3,-2);
    \draw[dotted] (3.5,-2) -- (4.5,-2);
    \draw (5,-2) -- (6,-2);
    \draw (7,-2) -- (8,-2);
    \draw (7,-2) -- (8,-2) node[right] {$s$};
    \draw[sim] (6.5,0.5) -- (7,0.5);
  \end{tikzpicture}.
\end{align}
The extra relative phase $e^{it\Delta_{\C,s}}$ on
$\tilde\beta_{sm}\p{t}$ can be cancelled out by the application of a
magnetic field with $U_\B\p{t\Delta_{\C,s}}$. In the basis
$\ket{m}_s\equiv\ket{s,-I;m}$ for site $s$ and
$\ket{n}_\C\equiv\ket{\C,I-n}$ for site $\C$, the net effect of our
sequence is thus the controlled-phase gate
\begin{multline}
  U_{\P,s}^{(0)}\p{t}
  \equiv U_\B\p{t\Delta_{\C,s}}
  U_{\pi,s}^- U_{\t{hold}}\p{t} U_{\pi,s}^- \\
  = \sum_{m,n\in\set{0,1}} \exp\p{-itV_snm} \op{m}_s\otimes\op{n}_\C
  = \op{0}_s \otimes \1_\C + \op{1}_s \otimes
  \begin{pmatrix}
    1 & 0 \\ 0 & e^{-itV_s}
  \end{pmatrix}_\C.
  \label{eq:U_P_0}
\end{multline}
A similar controlled-phase gate $U_{\P,s}^{(1)}\p{t}$ can be realized
conditional on the presence of an atom on site $s$ with nuclear spin
$-I+1$ by using $U_{\pi,s}^+$ in place of $U_{\pi,s}^-$.  Note that if
the atom on site $s$ is in a nontrivial superposition of nuclear spin
states $\ket{-I}$ and $\ket{-I+1}$, there will be two atoms on site
$\C$ during the hold time of $U_{\P,s}^{(0)}\p{t}$ and
$U_{\P,s}^{(1)}\p{t}$, resulting in atom losses from
${}^3P_0$-${}^3P_0$ interactions.

As a final comment, we mention that if we are guaranteed a single atom
on e.g.  site $\L$ and no atom on site $\R$, we can use the
nuclear-spin selective $\L\leftrightarrow\R$ tunneling gates mentioned
at the end of section \ref{sec:tunneling} to construct a
controlled-phase gate conditional on the nuclear spin of an atom on
site $s$, rather than on the mere presence of such an atom.  That is,
we can construct a gate similar to that in \eqref{eq:U_P_0}, but with
the nuclear spin basis $\set{\ket{-I},\ket{-I+1}}$ for site $s$.  We
can provide and explain this construction in more detail upon request;
the basic idea is to
\begin{enumerate*}[label=(\roman*)]
\item move e.g. the $-I+1$ spin component of the electronic
  ground-state atom in site $\L$ to site $\R$,
\item act with the controlled-phase gates $U_{\P,s}^{(0)}$ in
  \eqref{eq:U_P_0}, and then
\item move the $-I+1$ spin component of the electronic ground-state
  atom back to site $\L$.
\end{enumerate*}

\end{document}

### 2018-07-09 soc_squeezing/soc_squeezing_notes.tex
clarified some continuous two-axis twisting protocol and fixed error in OAT parameter table

@@
 squeezing time
-$t_{\t{opt}}^{\t{OAT}}\approx1.1N^{-2/3}\chi^{-1}\sim\eta
+$t_{\t{opt}}^{\t{OAT}}\sim N^{-2/3}\chi^{-1}\sim\eta
 N^{1/3}=N^{4/3}/L$.  Lattice depths are provided in units of the
@@
     Parameter & Symbol & Value \\ \hline\hline
-    Primary lattice depth & $V_0$ & 3 $E_R$ \\
+    Primary lattice depth & $V_0$ & 4 $E_R$ \\
     Transverse lattice depths & $V_T$ & 80 $E_R$ \\
@@
 \item modulate the drive as $\Omega\p{t}=\beta\omega\cos\p{\omega t}$
-  with $\omega\gg N\chi$,
-\item move into the rotating frame of $-\Omega\p{t}S_x$, and finally
-\item shift $H_{\t{drive,eff}}$ by the constant
-  $\p{\chi/2}\sp{\J_0\p{2\beta}-1} S^2$, where $\J_0\p{x}$ is the
-  zero-order Bessel function of the first kind evaluated at $x$.
+  with $\omega\gg N\chi$, and
+\item move into the rotating frame of $-\Omega\p{t}S_x$.
+  % , and finally
+% \item shift $H_{\t{drive,eff}}$ by the constant
+  % $\p{\chi/2}\sp{\J_0\p{2\beta}-1} S^2$, where $\J_0\p{x}$ is the
+  % zero-order Bessel function of the first kind evaluated at $x$.
 \end{enumerate*}
@@
   H_{\t{drive,eff}}^{\t{mod}}
-  = \f{\chi}{2} \sp{3\J_0\p{2\beta}-1} S_z^2
-  + \f{\chi}{2} \sp{1-\J_0\p{2\beta}} \p{S_z^2 - S_x^2},
+  = \f{\chi}{2} \p{\sp{\J_0\p{2\beta}+1} S_z^2
+    - \sp{\J_0\p{2\beta}-1} S_y^2},
 \end{align}
-where choosing the modulation index $\beta$ such that
-$\J_0\p{2\beta}=1/3$ or $\J_0\p{2\beta}=-1/3$ respectively results in
-the two-axis twisting Hamiltonians
+where $\J_0$ is the zero-order Bessel function of the first kind.
+Choosing a modulation index $\beta$ such that $\J_0\p{2\beta}=1/3$ or
+$\J_0\p{2\beta}=-1/3$ then respectively results in the two-axis
+twisting Hamiltonians
 \begin{align}
-  H_{\t{TAT}}^{z,x} = \f{\chi}{3} \p{S_z^2 - S_x^2},
+  H_{\t{TAT}}^{z,x}
+  = \f{\chi}{3} \p{2 S_z^2 + S_y^2}
+  \simeq \f{\chi}{3} \p{S_z^2 - S_x^2},
   &&
-  H_{\t{TAT}}^{y,x} = \f{\chi}{3}\p{S_y^2 - S_x^2}.
+  H_{\t{TAT}}^{y,x}
+  = \f{\chi}{3} \p{S_z^2 + 2 S_y^2}
+  \simeq \f{\chi}{3}\p{S_y^2 - S_x^2},
   \label{eq:H_TAT_drive}
 \end{align}
-Given the set of parameters in Table \ref{tab:parameters}, the
-two-axis twisting Hamiltonian $H_{\t{TAT}}^{z,x}$ can be achieved with
-$\omega=20\times2\pi$ Hz and $\beta=0.9057$.  The optimal squeezing
+where $\simeq$ denotes equality up to a global energy shift. Given the
+set of parameters in Table \ref{tab:parameters}, the two-axis twisting
+Hamiltonian $H_{\t{TAT}}^{z,x}$ can be achieved with
+$\omega=20\times2\pi$ Hz and $\beta=0.906$.  The optimal squeezing
 time for this protocol is $t_{\t{opt}}^{\t{TAT}}\approx0.98$ seconds.

### 2018-07-09 soc_squeezing/soc_squeezing_notes.tex
removed mean-field treatment changed figure directory from 'plots' to 'figures' added OAT and TAT comparison figure

@@
 \usepackage[caption=false]{subfig} % subfigures (via \subfloat[]{})
-\graphicspath{{./plots/}} % set path for all figures
+\graphicspath{{./figures/}} % set path for all figures
 
@@
 
-\section{Mean-field treatment}
-
-To sum up our developments thus far, spin-orbit coupling gives rise to
-an effective inhomogeneous field, while $s$-wave interactions give
-rise to a collective spin interaction, which means that the system is
-characterized by the Hamiltonian \note{[MP: I removed the mention of
-  non-uniform coupling (i.e.~through $\xi_{ij}$), as I feel it is
-  unnecessary given the preceding section on the collective spin
-  model.  The uniform all-to-all coupling in \eqref{eq:H_int} is exact
-  under the assumptions of the spin model, so there is no need to
-  ``re-derive'' it by throwing out inhomogeneous couplings and taking
-  $\xi_{ij}\to\xi\equiv\overline{\xi_{ij}}$.]}
-\begin{align}
-  H = H_{\t{int}} + H_{\t{SOC}} = -\xi\v S\c\v S - \sum_q h_q s_q^z,
-  \label{eq:H_full}
-\end{align}
-
-\note{[MP] Note: ``eqnarray'' environment is generally not recommended
-  for typesetting math, as it has inconsistent spacing in addition to
-  some other problems (discussed a bit
-  \href{http://www.tug.org/pracjourn/2006-4/madsen/madsen.pdf}{here},
-  if you're interested).  The ``align'' environment, on the other
-  hand, is provided by math packages for the purposes of typesetting
-  math, lacks spacing problems, and generally has more functionality.}
-
-This collective Heisenberg model is a celebrated model for magnetism,
-and is closely related to the Richardson-Gaudin models for
-superconductivity based on the Anderson pseudo-spin mapping
-\note{[cite?]}.  So before demonstrating how to generate an effective
-one-axis squeezing Hamiltonian based on Eqn.~\eqref{eq:H_full}, we
-first try to develop better understanding of this Hamiltonian.  In
-fact, this model is well-studied at the mean-field (MF) level as given
-by
-\begin{align}
-  H_{\t{MF}} = -\sum_q \v s_q \c \v B_q,
-  &&
-  \v B_q \equiv \xi\v{\mathcal S} + \v h_q,
-  &&
-  \v{\mathcal S} \equiv \bk{\v S},
-  &&
-  \v h_q \equiv \p{0,0,h_q}.
-  \label{eq:mean_field}
-\end{align}
-\note{[MP: I removed a factor of 2 in front of $\xi$, as I believe it
-  was a mistake.  Was this correct?  If the factor of 2 was indeed
-  incorrect, please make sure that there are no extra factors of 2 in
-  Peiru's codes.]}
-% MP: as an additional side note, I personally recommend the use of
-% \braket{X} or (\bk{X}, which I defined as a shorthand) in place of
-% \langle X \rangle; I find this is a little easier to read.
-% Similarly, you can use \ket{X} (\bra{X}) in place of |X\rangle in
-% place of (\langle X|), and \op{X}{Y} in place of |X\rangle\langle Y|
-% (where "op" stands for "outer product")
-
-The behavior of this model is determined by a competition between the
-non-local MF coupling through $\xi$ and an effective inhomogeneous
-axial field $\v h_q$ whose strength has a variance
-$8J_0^2\sin^2\p{\phi/2}$.  Fig.~\ref{fig:MFcontrast} shows the
-dynamics of the system as computed by numerical simulation of
-\eqref{eq:mean_field}, after initialization with all spins pointing
-along the $x$ axis, i.e.~in the state
-$\bigotimes_q\p{\ket{q,\g}+\ket{q,\e}}/\sqrt2$. \note{[MP: I added
-  text specifying the initial state; is it correct?]}  As the
-interaction strength $\xi$ increases, the contrast
-$\mathcal C\equiv2\abs{\mathcal S_+}/N$, \note{[MP: I propose writing
-  $\mathcal C\equiv\sqrt{\mathcal S_x^2+\mathcal S_y^2}/\p{N/2}$;
-  while longer, I feel like it makes the physical interpretation of
-  $\mathcal C$ more clear.]} i.e.~the normalized magnitude of the
-collective spin vector in the $x$-$y$ plane, gets stabilized against
-dephasing induced by the effective inhomogeneous axial field $\v h_q$.
-Furthermore, when $\xi$ is below a critical value $\xi_{\t{c}}$, the
-contrast exhibits some oscillatory behavior, but ultimately decays at
-long times.  Above the critical value $\xi_{\t{c}}$, however, the
-contrast reaches a non-zero steady state, with oscillatory behavior
-that is damped with increasing $\xi$.  These two phases can be
-understood through the underlying integrability of
-\eqref{eq:mean_field}.  Using the techniques in \note{[cite]}, the
-steady state contrast $\bar{\mathcal C}$ at the MF level can be
-obtained analytically as
-\begin{align}
-  \bar{\mathcal C} =
-  \begin{cases}
-    \sqrt{1-\sp{4J_0\sin\p{\phi/2}/N\xi}^2},
-    & \text{if}~ \abs{4J_0\sin(\phi/2)/N\xi} < 1 \\
-    0,
-    & \text{otherwise}
-  \end{cases}
-\end{align}
-
-\begin{figure}
-\centering
-\includegraphics[width=3.5in]{MFcontrast.pdf}
-\caption{
-The contrast as a function of time and interaction strength.
-}
-\label{fig:MFcontrast}
-\end{figure}
-
-Comparing the MF simulation results with the quantum simulation
-results \note{[MP: In my experience, ``quantum simulation'' refers to
-  that which is performed with a quantum simulator, i.e.~via
-  experiment.  I assume that's not what you meant here, and that these
-  are actually some sort of Monte Carlo / quantum trajectory / TWA
-  simulations.  Can we specify the method?]}, however, we find that
-the contrast decreases rather than remain as a constant even if
-$\xi\gg\xi_{\t{c}}$, as shown in Fig.~\ref{fig:EffectiveSqueezing}.
-
-\begin{figure}
-\centering
-\includegraphics[width=3in]{Fig33.pdf}
-\caption{Contrast comparison among the MF simulation results, quantum simulation results and analytic expressions:
-(a) Good agreement between MF simulation (solid line) and quantum simulation (dashed line) at short time,
-while  the discrepancy occurs at longer time due to the quantum fluctuations.
-(b-c) Benchmark the approximated one-axis twisting analytical expressions (dashed line) with the quantum treatment (solid line).
-}
-\label{fig:EffectiveSqueezing}
-\end{figure}
-
-\note{[MP: While it may be nice for understanding some properties of
-  the Hamiltonian in \eqref{eq:H_full}, I do not think that this text
-  about contrast and mean-field treatment adds much to our goal of
-  spin squeezing.  Do we need to include this material here?  I would
-  cut this section out for now.]}
-
 \section{Effective spin squeezing}
@@
 \item move into the rotating frame of $-\Omega\p{t}S_x$.
-  % , and finally
-% \item shift $H_{\t{drive,eff}}$ by the constant
-  % $\p{\chi/2}\sp{\J_0\p{2\beta}-1} S^2$, where $\J_0\p{x}$ is the
-  % zero-order Bessel function of the first kind evaluated at $x$.
 \end{enumerate*}
@@
 time for this protocol is $t_{\t{opt}}^{\t{TAT}}\approx0.98$ seconds.
+A comparison of squeezing through $H_{\t{OAT}}$ and
+$H_{\t{TAT}}^{z,x}$ via the parameters in Table \ref{tab:parameters}
+is provided in Figure \ref{fig:squeezing_comparison}.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics{squeezing_comparison.pdf}
+  \caption{Comparison of squeezing through $H_{\t{OAT}}$ and
+    $H_{\t{TAT}}^{z,x}$ via the parameters in Table
+    \ref{tab:parameters} and the continuous two-axis twisting protocol
+    with $\omega=20\times2\pi$ Hz and $\beta=0.906$.}
+  \label{fig:parameters}
+\end{figure}
 

### 2018-07-10 soc_squeezing/soc_squeezing_notes.tex
fixed misnamed figure
tweaked example parameters for OAT and TAT

@@
     Primary lattice depth & $V_0$ & 4 $E_R$ \\
-    Transverse lattice depths & $V_T$ & 80 $E_R$ \\
+    Transverse lattice depths & $V_T$ & 60 $E_R$ \\
     Lattice sites & $L$ & 100 \\
@@
 Hamiltonian $H_{\t{TAT}}^{z,x}$ can be achieved with
-$\omega=20\times2\pi$ Hz and $\beta=0.906$.  The optimal squeezing
-time for this protocol is $t_{\t{opt}}^{\t{TAT}}\approx0.98$ seconds.
-A comparison of squeezing through $H_{\t{OAT}}$ and
-$H_{\t{TAT}}^{z,x}$ via the parameters in Table \ref{tab:parameters}
-is provided in Figure \ref{fig:squeezing_comparison}.
+$\omega=N\abs{\xi\chi}^{1/2}\approx27\times2\pi$ Hz and $\beta=0.906$.
+The optimal squeezing time for this protocol is
+$t_{\t{opt}}^{\t{TAT}}\approx0.98$ seconds.  A comparison of squeezing
+through $H_{\t{OAT}}$ and $H_{\t{TAT}}^{z,x}$ via the parameters in
+Table \ref{tab:parameters} is provided in Figure
+\ref{fig:squeezing_comparison}.
 
@@
     \ref{tab:parameters} and the continuous two-axis twisting protocol
-    with $\omega=20\times2\pi$ Hz and $\beta=0.906$.}
-  \label{fig:parameters}
+    with $\omega=27\times2\pi$ Hz and $\beta=0.906$.}
+  \label{fig:squeezing_comparison}
 \end{figure}

### 2018-07-18 soc_squeezing/soc_squeezing_notes.tex
changed \xi to U/L in spin parameter, as \xi is already used for squeezing fixed previously incorrect optimal squeezing times

@@
   U \equiv G \int \d^3x~ \abs{w_0\p{x}}^4,
-  &&
-  \xi \equiv \f{U}{L},
 \end{align}
@@
 \begin{align}
-  H_{\t{int}} = \xi \sum_{p,q,r,s}
+  H_{\t{int}} = \f{U}{L} \sum_{p,q,r,s}
   \delta_{p+q,r+s} \hat c_{r,\e}^\dag \hat c_{s,\g}^\dag
@@
   H_{\t{int}}
-  = \xi \sum_{p,q}
+  = \f{U}{L} \sum_{p,q}
   \p{\hat c_{p,\e}^\dag \hat c_{q,\g}^\dag \hat c_{q,\g} \hat c_{p,\e}
     + \hat c_{q,\e}^\dag \hat c_{p,\g}^\dag \hat c_{q,\g} \hat c_{p,\e}}
-  = \xi \sum_{p,q}
+  = \f{U}{L} \sum_{p,q}
   \p{\hat c_{q,\g}^\dag \hat c_{q,\g} \hat c_{p,\e}^\dag \hat c_{p,\e}
@@
 \begin{align}
-  H_{\t{int}} = \xi \sum_{p,q}
+  H_{\t{int}} = \f{U}{L} \sum_{p,q}
   \sp{\p{\f{\1_q-\sigma_q^z}{2}} \p{\f{\1_p+\sigma_p^z}{2}}
@@
   H_{\t{int}}
-  = - \f14 \xi \sum_{p,q} \v{\sigma}_q\c\v{\sigma}_p
-  = - \xi \sum_{p,q} \v s_q \c\v s_p
-  = - \xi \v S \c \v S,
+  = - \f14 \f{U}{L} \sum_{p,q} \v{\sigma}_q\c\v{\sigma}_p
+  = - \f{U}{L} \sum_{p,q} \v s_q \c\v s_p
+  = - \f{U}{L} \v S \c \v S,
 \end{align}
@@
 (pseudo-)spin $S$ and projection $m$ onto a quantization axis; the
-corresponding energies are $\bk{Sm|H_{\t{int}}|S,m}=-\xi S\p{S+1}$.
+corresponding energies are $\bk{Sm|H_{\t{int}}|S,m}=-\p{U/L}S\p{S+1}$.
 The ground-state $S=N/2$ manifold is spanned by the Dicke states
@@
 electronic state excitations in the state $\ket{m}$.  Above this
-ground-state manifold with an energy gap of $N\xi$ lies the manifold
-of $S=N/2-1$ states, which is spanned by the orthonormal spin-wave
-states
+ground-state manifold with an energy gap of $NU/L=\eta U$ lies the
+manifold of $S=N/2-1$ states, which is spanned by the orthonormal
+spin-wave states
 \begin{align}
@@
 $s_n^+\equiv\hat c_{n,\e}\hat c_{n,\g}$
-\cite{swallows2011suppression}.  \note{[We should comment about how we
-  do not actually achieve $J_0\gg U$ in practice, but that we can
-  nonetheless validate the spin model using numerics.]}
+\cite{swallows2011suppression}.  \note{[We should comment about or
+  otherwise modify the text to reflect the fact that do not actually
+  achieve $J_0\gg U$ in practice.]}
 
@@
   H_{\t{free}} = H_{\t{int}} + H_{\t{SOC}}
-  = -\xi\v S\c\v S - \sum_n h_n s_n^z
-  = -\xi\v S\c\v S - \bar h S_z - \sum_n b_n s_n^z,
+  = -\f{U}{L} \v S\c\v S - \sum_n h_n s_n^z
+  = -\f{U}{L} \v S\c\v S - \bar h S_z - \sum_n b_n s_n^z,
 \end{align}
@@
 small compared to the collective spin gap, i.e.~when
-$\max_n\abs{b_n}\ll N\abs{\xi}$.  Such a perturbative treatment yields
-an effective Hamiltonian of the form
+$\max_n\abs{b_n}\ll \eta\abs{U}$.  Such a perturbative treatment
+yields an effective Hamiltonian of the form
 $H_{\t{free,eff}} = H_{\t{free,eff}}^{(1)} + H_{\t{free,eff}}^{(2)}$
@@
   &= \sum_m \op{m} H \op{m}
-  = -\xi \v S \c \v S - \bar h S_z, \\
+  = -\f{U}{L} \v S \c \v S - \bar h S_z, \\
   H_{\t{free,eff}}^{(2)}
@@
   = -\sum_{m,k}
-  \frac{\abs{\sum_n b_n \bk{mk|s_n^z|m}}^2}{N\xi} \op{m}
-  = -\sum_m \f{(N/2)^2-m^2}{N^3(N-1)\xi}
+  \frac{\abs{\sum_n b_n \bk{mk|s_n^z|m}}^2}{\eta U} \op{m}
+  = -\sum_m \f{(N/2)^2-m^2}{N^2\p{N-1}\eta U}
   \sum_k \abs{\sum_n b_n e^{2\pi ikn/N}}^2 \op{m}.
@@
   H_{\t{free,eff}}^{(2)}
-  = -\sum_m \f{\p{N/2}^2-m^2}{N\p{N-1}\xi}~ \tilde h^2 \op{m},
+  = -\sum_m \f{\p{N/2}^2-m^2}{\p{N-1}\eta U}~ \tilde h^2 \op{m},
 \end{align}
@@
   H_{\t{free,eff}}
-  = -\xi \v S \c \v S - \bar h S_z + \chi S_z^2,
+  = -\f{U}{L} \v S \c \v S - \bar h S_z + \chi S_z^2,
   &&
-  \chi \equiv \f{\tilde h^2}{N\p{N-1}\xi}.
+  \chi \equiv \f{\tilde h^2}{\p{N-1}\eta U}.
 \end{align}
@@
 \begin{align}
-  H_{\t{OAT}} = \chi S_z^2.
+  H_{\t{OAT}}^z = \chi S_z^2.
   \label{eq:H_OAT}
@@
 An example set of parameters and values for one-axis twisting via
-$H_{\t{OAT}}$ with strontium-87 is provided in Table
+$H_{\t{OAT}}^z$ with strontium-87 is provided in Table
 \ref{tab:parameters}, along with the optimal one-axis twisting
@@
     On-site interaction strength & $U$ & $\approx1.4\times2\pi$ kHz \\
-    Collective interaction strength & $\xi$
-    & $\approx14\times2\pi$ Hz \\
     Squeezing strength & $\chi$ & $\approx21\times2\pi$ mHz \\
     Optimal squeezing time & $t_{\t{opt}}^{\t{OAT}}$
-    & $\approx0.66$ seconds \\ \hline
+    & $\approx0.59$ seconds \\ \hline
   \end{tabular}
@@
 can neglect $H_{\t{eff}}$) to realize collective rotations of the form
-$\exp\p{\pm i\theta S_x}$.  Thus acting with $\exp\sp{i\p{\pi/2}S_x}$,
-waiting for a time $2\tau$, acting with $\exp\sp{-i\p{\pi/2}S_x}$, and
-waiting for a time $\tau$ then realizes the unitary
+$\exp\p{\pm i\theta S_x}$.  Thus
+\begin{enumerate*}[label=(\roman*)]
+\item acting with $\exp\sp{i\p{\pi/2}S_x}$,
+\item waiting for a time $2\tau$,
+\item acting with $\exp\sp{-i\p{\pi/2}S_x}$, and
+\item waiting for a time $\tau$
+\end{enumerate*}
+realizes the unitary
 \begin{align}
@@
 above because $S^2$ is conserved by all dynamics.  Repeating the above
-sequence $n$ times for a time $t=3n\tau$ thus results in the two-axis
-twisting unitary
+sequence $n$ times for a time $t=3n\tau$ thus results in the unitary
 \begin{align}
@@
 \end{align}
-which is equivalent to evolving for a time $t$ under the two-axis
-twisting Hamiltonian
+which is equivalently generated by evolution for a time $t$ under the
+two-axis twisting Hamiltonian
 \begin{align}
@@
 the optimal squeezing time for this two-axis twisting Hamiltonian is
-generally close to that for one-axis twisting with $H_{\t{OAT}}$ in
+generally close to that for one-axis twisting with $H_{\t{OAT}}^z$ in
 \eqref{eq:H_OAT} \cite{liu2011spin}.
@@
   simulation, we can probably approximate these gates by square pulses
-  centered on the appropriate time.  I believe that experiments are
-  capable of achieving $\Omega\sim1\times2\pi~\t{kHz}$.  I can do an
-  error analysis on this sequence to determine the effect of the
-  $\O\p{\tau^2\chi^2}$ terms in the exponent of $U_1$, but it is
-  probably simpler to just try this out in simulations.]}
+  centered on the appropriate time.  I can do an error analysis on
+  this sequence to determine the effect of the $\O\p{\tau^2\chi^2}$
+  terms in the exponent of $U_1$, but it is probably simpler to just
+  try this out in simulations.]}
+
+\note{[MP: I expect this pulsed-drive protocol to be worse than the
+  continuous-drive protocol below.  If we decide that we are not
+  interested in the pulsed drive, we can merely mention it and throw
+  in a citation of ref.~\cite{liu2011spin} without devoting an entire
+  section to the subject.]}
 
@@
   H_{\t{drive}} = H_{\t{free}} + H_{\t{clock}}
-  = -\xi\v S\c\v S - \bar h S_z - \Omega S_x - \sum_n b_n s_n^z,
+  = -\f{U}{L} \v S\c\v S - \bar h S_z - \Omega S_x - \sum_n b_n s_n^z,
 \end{align}
@@
 shown in Appendix \ref{sec:effective_squeezing_drive}, the addition of
-a drive with $\abs{\Omega}\ll N\abs{\xi}$ does not modify the
-squeezing behavior of the spin-orbit coupling terms in
+a drive with $\abs{\Omega}\ll\eta\abs{U}=N\abs{U}/L$ does not modify
+the squeezing behavior of the spin-orbit coupling terms in
 $H_{\t{drive}}$, whose effective action on the $S=N/2$ manifold is
@@
 set of parameters in Table \ref{tab:parameters}, the two-axis twisting
-Hamiltonian $H_{\t{TAT}}^{z,x}$ can be achieved with
-$\omega=N\abs{\xi\chi}^{1/2}\approx27\times2\pi$ Hz and $\beta=0.906$.
-The optimal squeezing time for this protocol is
-$t_{\t{opt}}^{\t{TAT}}\approx0.98$ seconds.  A comparison of squeezing
-through $H_{\t{OAT}}$ and $H_{\t{TAT}}^{z,x}$ via the parameters in
-Table \ref{tab:parameters} is provided in Figure
-\ref{fig:squeezing_comparison}.
+Hamiltonian $H_{\t{TAT}}^{z,x}$ can be achieved with a drive frequency
+$\omega=N\abs{\chi U/L}^{1/2}\approx27\times2\pi$ Hz and modulation
+index $\beta=0.906$.  The optimal squeezing time for this protocol is
+$t_{\t{opt}}^{\t{TAT}}\approx0.82$ seconds.  A comparison of squeezing
+through $H_{\t{OAT}}^z$ and $H_{\t{TAT}}^{z,x}$ for the initial state
+$\ket{Y}\equiv\bigotimes_n\p{\ket{\g}_n+i\ket{\e}_n}/\sqrt2$ via the
+parameters in Table \ref{tab:parameters} is provided in Figure
+\ref{fig:squeezing_comparison}.  The figure of merit for squeezing is
+the normalized minimal variance of spin in the plane orthogonal to the
+mean spin vector $\bk{\v S}$:
+\begin{align}
+  \xi^2 \equiv \f{N}{\abs{\bk{\v S}}^2}
+  \min_{\hat{\v n}\perp\bk{\v S}} \bk{\p{\v S\c\hat{\v n}}^2}
+\end{align}
+for $\abs{\hat{\v n}}=1$.
 
-\begin{figure}[h]
+\begin{figure}
   \centering
   \includegraphics{squeezing_comparison.pdf}
-  \caption{Comparison of squeezing through $H_{\t{OAT}}$ and
+  \caption{Comparison of squeezing through $H_{\t{OAT}}^z$ and
     $H_{\t{TAT}}^{z,x}$ via the parameters in Table

### 2018-07-19 soc_squeezing/soc_squeezing_notes.tex
moved derivation of OAT Hamiltonian to an appendix

@@
 \usepackage{dsfont} % for identity operator
-\newcommand{\1}{\hat{\mathds{1}}}
+\newcommand{\1}{\mathds{1}}
 
@@
 
-These are some notes about trying to induce squeezing in an optical
-lattice clock by combining spin-orbit coupling and interactions.  We
-assume, for now, that
+These are some notes about trying to induce spin squeezing in an
+optical lattice clock by combining spin-orbit coupling, interactions,
+and external drives.  We assume, for now, that
 \begin{enumerate*}[label=(\roman*)]
@@
 
-If the single-particle energy spacings are much greater than the
+If the single-particle energy spacings are large compared to the
 strength of inter-particle interactions, or equivalently if
-$J_0\gg U$, then by the secular approximation we can neglect terms in
-$H_{\t{int}}$ which do not conserve the sum of single-particle
+$J_0\gtrsim U$, then by the secular approximation we can neglect terms
+in $H_{\t{int}}$ which do not conserve the sum of single-particle
 energies.  By solving this single-particle energy conservation
@@
 corresponding energies are $\bk{Sm|H_{\t{int}}|S,m}=-\p{U/L}S\p{S+1}$.
-The ground-state $S=N/2$ manifold is spanned by the Dicke states
+In particular, the ground-state $S=N/2$ manifold is spanned by the
+Dicke states
 \begin{align}
@@
   S_+ \equiv \sum_n \sigma_n^+,
+  \label{eq:dicke_states}
 \end{align}
-for which we note that $N/2+m=0,1,\cdots,N$ is the number of
-electronic state excitations in the state $\ket{m}$.  Above this
-ground-state manifold with an energy gap of $NU/L=\eta U$ lies the
-manifold of $S=N/2-1$ states, which is spanned by the orthonormal
-spin-wave states
-\begin{align}
-  \ket{mk}
-  \equiv \ket{N/2-1,m,k}
-  \equiv {N-1 \choose \p{N/2-m}\p{N/2-m+1}}^{-1/2}
-  \sum_{n=1}^N e^{2\pi ikn/N} s_n^+ \ket{N/2,m-1},
-\end{align}
-where $k=1,2,\cdots,N-1$; $n$ indices an individual atom; and
-$s_n^+\equiv\hat c_{n,\e}\hat c_{n,\g}$
-\cite{swallows2011suppression}.  \note{[We should comment about or
-  otherwise modify the text to reflect the fact that do not actually
-  achieve $J_0\gg U$ in practice.]}
+where $n$ indexes an individual atom,
+$\sigma_n^+\equiv\hat c_{n,\e}\hat c_{n,\g}$ is an individual
+spin-raising operator, and $N/2+m=0,1,\cdots,N$ is the number of
+electronic state excitations in the state $\ket{m}$
+\cite{swallows2011suppression}.  In words, the Dicke state $\ket{m}$
+is a uniform superposition of all states with $N/2+m$ total electronic
+excitations.
 
@@
 
-In order to demonstrate spin squeezing behavior induced by spin-orbit
-coupling, we fix the total particle number $N$, consider initial
-states with total spin $S=N/2$, and write the total system Hamiltonian
-as
+In total, free evolution of atoms on a lattice is governed by the
+Hamiltonian
 \begin{align}
-  H_{\t{free}} = H_{\t{int}} + H_{\t{SOC}}
-  = -\f{U}{L} \v S\c\v S - \sum_n h_n s_n^z
-  = -\f{U}{L} \v S\c\v S - \bar h S_z - \sum_n b_n s_n^z,
+  H_{\t{free}}
+  = H_{\t{int}} + H_{\t{SOC}}
+  = -\f{U}{L} \v S\c\v S - \sum_n h_n s_n^z,
 \end{align}
-where we have separated out the mean and residual effective fields
+for $n$ indexing an individual atom.  When the ``local'' fields $h_n$
+are small in magnitude compared to the collective spin gap $\eta U$
+and we initialize all atoms in the ground-state subspace of
+$H_{\t{int}}$, we can treat the action of the spin-orbit coupling
+Hamiltonian perturbatively, which yields the effective free-evolution
+Hamiltonian (see Appendix \ref{sec:squeezing_derivation} or Peiru's
+notes)
+\begin{align}
+  H_{\t{free,eff}} = -\bar h S_z + \chi S_z^2,
+\end{align}
+where
 \begin{align}
@@
   &&
-  b_n \equiv h_n - \bar h.
-\end{align}
-The subscript on $H_{\t{free}}$ refers to the fact that this
-Hamiltonian corresponds to free, undriven evolution of (interacting)
-atoms on the lattice.  The collective spin operators $\v S\c\v S$ and
-$S_z$ both preserve total spin $S$ and its projection $m$ onto a
-quantization axis, which makes the Dicke states a good basis for
-treating $H_{\t{free}}$ (i.e.~within the $S=N/2$ manifold).  While the
-residual effective field term $\sum_n b_n s_n^z$ generally takes
-states outside the $S=N/2$ manifold, we can treat this spin-changing
-process perturbatively when the residual effective fields $b_n$ are
-small compared to the collective spin gap, i.e.~when
-$\max_n\abs{b_n}\ll \eta\abs{U}$.  Such a perturbative treatment
-yields an effective Hamiltonian of the form
-$H_{\t{free,eff}} = H_{\t{free,eff}}^{(1)} + H_{\t{free,eff}}^{(2)}$
-with
-\begin{align}
-  H_{\t{free,eff}}^{(1)}
-  &= \sum_m \op{m} H \op{m}
-  = -\f{U}{L} \v S \c \v S - \bar h S_z, \\
-  H_{\t{free,eff}}^{(2)}
-  &= -\sum_{\substack{m\\S<N/2}}
-  \frac{\abs{\bk{Sm|H|m}}^2}{E_{Sm}^{\t{free}}-E_m^{\t{free}}} \op{m}
-  = -\sum_{\substack{m\\S<N/2}}
-  \frac{\abs{\sum_n b_n \bk{Sm|s_n^z|m}}^2}
-  {E_{Sm}^{\t{free}}-E_m^{\t{free}}} \op{m},
-\end{align}
-where $E_\psi^{\t{free}}\equiv\bk{\psi|H_{\t{free}}|\psi}$ and we have
-used the fact the spin projection $m$ is conserved by $H$.  As the
-individual spin operators $s_n^z$ cannot change the total spin $S$ by
-more than 1, we need only consider $S=N/2-1$ spin-wave states when
-evaluating the sum over $S<N/2$ in $H_{\t{free,eff}}^{(2)}$ above.
-These spin-wave states couple to the Dicke states via the matrix
-elements of the form (see section 2 in the supporting materials of
-\cite{swallows2011suppression})
-\begin{align}
-  \bk{m|s_n^z|mk} = e^{2\pi i k n/N} \sqrt{\f{(N/2)^2-m^2}{N^2 (N-1)}},
-\end{align}
-which implies
-\begin{align}
-  H_{\t{free,eff}}^{(2)}
-  = -\sum_{m,k}
-  \frac{\abs{\sum_n b_n \bk{mk|s_n^z|m}}^2}{\eta U} \op{m}
-  = -\sum_m \f{(N/2)^2-m^2}{N^2\p{N-1}\eta U}
-  \sum_k \abs{\sum_n b_n e^{2\pi ikn/N}}^2 \op{m}.
-\end{align}
-Using the fact that $\sum_nb_n=0$, we can expand
-\begin{align}
-  \sum_k \abs{\sum_n b_n e^{2\pi ikn/N}}^2
-  = \sum_{k=1}^{N-1} \sum_{n,n'} b_n b_{n'} e^{2\pi ik\p{n-n'}/N}
-  = \sum_{n,n'} b_n b_{n'} \sum_{k=0}^{N-1} e^{2\pi ik\p{n-n'}/N},
-\end{align}
-where the sum over $k$ vanishes for $n\ne n'$ and equals $N$ when
-$n=n'$, so
-\begin{align}
-  \sum_k \abs{\sum_n b_n e^{2\pi ikn/N}}^2
-  = N \sum_n b_n^2 = N^2 \tilde h^2,
-  &&
-  \tilde h^2 \equiv \f1N \sum_n b_n^2 = \f1N \sum_n \p{h_n - \bar h}^2.
-  \label{eq:sum_knn}
-\end{align}
-We therefore have that
-\begin{align}
-  H_{\t{free,eff}}^{(2)}
-  = -\sum_m \f{\p{N/2}^2-m^2}{\p{N-1}\eta U}~ \tilde h^2 \op{m},
-\end{align}
-where the term $\p{N/2}^2$ term contributes a global energy shift
-which we can neglect, while the $m^2$ term is proportional to
-$m^2\op{m}=S_z^2$.  In total, the effective Hamiltonian is thus
-\begin{align}
-  H_{\t{free,eff}}
-  = -\f{U}{L} \v S \c \v S - \bar h S_z + \chi S_z^2,
+  \tilde h^2 \equiv \f1N \sum_n \p{h_n - \bar h}^2,
   &&
@@
 \end{align}
-This effective Hamiltonian preserves total spin $S$, which means that
-the $\v S\c\v S$ term contributes only a constant energy shift which
-we can neglect.  Furthermore, at temperatures $T\gg J_0$ the mean
-effective field $\bar h=0$, while for $T\sim J_0$ the collective
-rotation due to the $S_z$ term can be neglected by use of a rotating
-frame.  The effective Hamiltonian therefore simplifies to the one-axis
-twisting Hamiltonian
+At temperatures $T\gg J_0$, the mean effective field $\bar h=0$, while
+for $T\sim J_0$ the collective rotation due to the $S_z$ term can be
+eliminated by use of a rotating frame.  The free-evolution Hamiltonian
+thus simplifies further to the one-axis twisting Hamiltonian
 \begin{align}
@@
 An example set of parameters and values for one-axis twisting via
-$H_{\t{OAT}}^z$ with strontium-87 is provided in Table
-\ref{tab:parameters}, along with the optimal one-axis twisting
-squeezing time
+$H_{\t{OAT}}^z$ with strontium-87 in a 1-D magic-wavelength lattice is
+provided in Table \ref{tab:parameters}, along with the optimal
+one-axis twisting squeezing time
 $t_{\t{opt}}^{\t{OAT}}\sim N^{-2/3}\chi^{-1}\sim\eta
@@
 which is equivalently generated by evolution for a time $t$ under the
-two-axis twisting Hamiltonian
+two-axis countertwisting Hamiltonian
 \begin{align}
@@
 \end{align}
-Despite the factor of 1/3 in the effective two-axis twisting strength,
-the optimal squeezing time for this two-axis twisting Hamiltonian is
-generally close to that for one-axis twisting with $H_{\t{OAT}}^z$ in
-\eqref{eq:H_OAT} \cite{liu2011spin}.
+Despite the factor of 1/3 in the effective two-axis countertwisting
+strength, the optimal squeezing time for this two-axis twisting
+Hamiltonian is generally close to that for one-axis twisting with
+$H_{\t{OAT}}^z$ in \eqref{eq:H_OAT} \cite{liu2011spin}.
 
@@
 \begin{figure}
-  \centering
-  \includegraphics{squeezing_comparison.pdf}
+  \centering \includegraphics{squeezing_comparison_TAT.pdf}
   \caption{Comparison of squeezing through $H_{\t{OAT}}^z$ and
@@
 
-\section{Effective spin squeezing in the presence of a drive}
-\label{sec:effective_squeezing_drive}
+\section{Effective spin squeezing in the presence of a weak drive}
+\label{sec:squeezing_derivation}
 
@@
   &&
-  H_0 = - \xi\v S\c\v S,
+  H_0 = - \f{U}{L} \v S\c\v S,
   &&
-  V = - \bar h S_z - \Omega S_x - \sum_n b_n s_n^z,
+  V = - \sum_n h_n s_n^z - \Omega S_x,
 \end{align}
-where the effective fields $b_n$ are mean-zero, and we consider
-initial states in the ground-state manifold $\G_0$ of $H_0$ (i.e.~with
-total spin $S=N/2$).  If the largest eigenvalue of $V$ is smaller in
-magnitude than half of the collective spin gap $N\xi$, or the energy
-gap under $H_0$ between $\G_0$ and its orthogonal complement $\E_0$,
-then we can formally develop a perturbative treatment of the action of
-$V$ on $\G_0$.  Such a treatment yields an effective Hamiltonian
-$H_{\t{eff}}=\sum_pH_{\t{eff}}^{(p)}$ on $\G_0$, where
+and we consider $N$-particle states initially in the ground-state
+manifold $\G_0$ of $H_0$, which have total spin $S=N/2$.  If the
+largest eigenvalue of $V$ is smaller in magnitude than half of the
+collective spin gap $N\abs{U}/L=\eta\abs{U}$, i.e.~the energy gap
+under $H_0$ between $\G_0$ and its orthogonal complement $\E_0$, then
+we can formally develop a perturbative treatment for the action of $V$
+on $\G_0$.  Such a treatment yields an effective Hamiltonian on $\G_0$
+of the form $H_{\t{eff}}=\sum_pH_{\t{eff}}^{(p)}$, where
 $H_{\t{eff}}^{(p)}$ is order $p$ in $V$.  Letting $\P_0$ ($\Q_0$) be a
@@
 \end{align}
-The first few terms in the expansion of the effective Hamiltonian are
-then, as derived in ref.~\cite{bravyi2011schrieffer},
+The first few terms in the expansion of the effective Hamiltonian
+$H_{\t{eff}}$ are then, as derived in
+ref.~\cite{bravyi2011schrieffer},
 \begin{align}
@@
   H_{\t{eff}}^{(2)} = -\f12 \P_0 \sp{\O V,\L V} \P_0.
+  \label{eq:H_eff_012}
 \end{align}
-Letting $\B_0\p{\E_0}$ denote an eigenbasis of $H_0$ for $\E_0$ and
-adding a constant shift to $H_0$ such that $E_\psi=0$ for
-$\ket\psi\in\G_0$, we additionally define
+If we add a constant to $H_0$ such that $E_\psi=0$ for all
+$\ket\psi\in\G_0$, or equivalently we measure all energies relative to
+that of the ground-state manifold $\G_0$ with respect to $H_0$, it
+immediately follows that $H_{\t{eff}}^{(0)}=0$.  To calculate
+$H_{\t{eff}}^{(1)}$, we note that the ground-state manifold $\G_0$ is
+spanned by the Dicke states $\ket{m}$ defined in
+\eqref{eq:dicke_states}, in terms of which we can expand the
+collective spin-$z$ operator as $S_z=\sum_mm\op{m}$.  We can likewise
+expand the collective spin-$x$ operator $S_x$ in terms of $x$-oriented
+Dicke states $\ket{m_x}$ as $S_x=\sum_mm\op{m_x}$.  The ground-state
+projector $\P_0$ onto $\G_0$ can be expanded in either basis as
+$\P_0=\sum_m\op{m}=\sum_m\op{m_x}$.  Defining the mean and residual
+fields
+\begin{align}
+  \bar h \equiv \f1N \sum_n h_n,
+  &&
+  b_n \equiv h_n - \bar h,
+\end{align}
+we can then write
+\begin{align}
+  V = -\sum_n \p{b_n+\bar h} s_n^z - \Omega S_x
+  = -\sum_n b_n s_n^z - \bar h S_z - \Omega S_x,
+\end{align}
+and in turn
+\begin{align}
+  H_{\t{eff}}^{(1)}
+  = \P_0\p{-\sum_n b_n s_n^z - \bar h S_z - \Omega S_x} \P_0
+  = -\sum_n b_n \P_0 s_n^z\P_0 - \bar h S_z - \Omega S_x,
+\end{align}
+where we used the fact that $\P_0 S_{j=z,x} \P_0 = S_j$.  By
+construction, the residual fields are mean-zero, i.e.~$\sum_nb_n=0$.
+Using the particle-exchange symmetry of the Dicke states, we can thus
+expand
+\begin{align}
+  \sum_n b_n \P_0 s_n^z \P_0
+  = \sum_{n,m,m'} b_n \op{m} s_n^z \op{m'}
+  = \sum_n b_n \sum_{m,m'} \op{m} s_1^z \op{m'}
+  = 0,
+\end{align}
+which implies
+\begin{align}
+  H_{\t{eff}}^{(1)} = - \bar h S_z - \Omega S_x.
+\end{align}
+To calculate the second-order effective Hamiltonian
+$H_{\t{eff}}^{(2)}$, we let $\B_0\p{\E_0}$ denote an eigenbasis of
+$H_0$ for the excited subspace $\E_0$, define the operator
 \begin{align}
@@
 \end{align}
-The expression for $H_{\t{eff}}^{(2)}$ then simplifies to
+The expression for $H_{\t{eff}}^{(2)}$ in \eqref{eq:H_eff_012} then
+simplifies to
 \begin{align}
@@
 \end{align}
-\note{[MP: much of the text above is essentially copied from an
-  appendix of the theory paper for multi-body interactions.  As soon
-  as that paper gets posted on arXiv, we can cite it instead.]}  It
-follows that
+The only part of $V$ which is off-diagonal with respect to the ground-
+and excited-state manifolds $\G_0$ and $\E_0$ is $-\sum_nb_ns_n^z$,
+and the individual operators in this term can only change the total
+spin $S$ by at most 1.  It is therefore sufficient to expand $\I$ in a
+basis for states with total spin $S=N/2-1$, which is provided by the
+spin-wave states
 \begin{align}
-  H_{\t{eff}}^{(0)} = 0,
-  &&
-  H_{\t{eff}}^{(1)} = -\bar h S_z - \Omega S_x,
+  \ket{mk}
+  \equiv \ket{N/2-1,m,k}
+  \equiv {N-1 \choose \p{N/2-m}\p{N/2-m+1}}^{-1/2}
+  \sum_{n=1}^N e^{2\pi ikn/N} s_n^+ \ket{N/2,m-1},
 \end{align}
-and, in terms of the $S=N/2$ Dicke states $\ket{m}\equiv\ket{N/2,m}$
-and the $S=N/2-1$ spin-wave states $\ket{mk}\equiv\ket{N/2-1,mk}$,
+for $k=1,2,\cdots,N-1$ \cite{swallows2011suppression}.  Using the fact that all spin-$z$ operators preserve the projection of total spin onto the $z$ axis, we then have that
 \begin{align}
   H_{\t{eff}}^{(2)}
-  = - \f1{N\xi} \sum_{m,k} \abs{\sum_nb_n\bk{mk|s_n^z|m}}^2 \op{m},
+  = -\f1{\eta U} \sum_{m,m',k,n,n'} b_n b_{n'}
+  \Bk{m| s_n^z \op{m'k} s_{n'}^z |m} \op{m}.
   \label{eq:general_H_eff_2}
 \end{align}
-where we used the fact that the collective spin operators $S_j$
-conserve total spin $S$, while the individual spin operators $s_n^z$
-conserve the projection $m$ of collective spin onto a quantization
-axis.  The second-order effective Hamiltonian $H_{\t{eff}}^{(2)}$ in
-\eqref{eq:general_H_eff_2} is identical to that derived in Section
-\ref{sec:OAT}, which implies
+The relevant matrix elements between the Dicke states and the
+spin-wave states are \cite{swallows2011suppression}
 \begin{align}
-  H_{\t{eff}}^{(2)} = \chi S_z^2,
+  \bk{m|s_n^z|m'k}
+  = e^{2\pi i k n/N} \sqrt{\f{(N/2)^2-m^2}{N^2 (N-1)}}~ \delta_{m,m'},
+\end{align}
+which implies
+\begin{align}
+  H_{\t{free,eff}}^{(2)}
+  = -\f1{\eta U} \sum_m \f{(N/2)^2-m^2}{N^2 (N-1)} \op{m}
+  \sum_{k,n,n'} b_n b_{n'} e^{2\pi ik\p{n-n'}/N}.
+\end{align}
+Using the fact that $\sum_nb_n=0$, we can expand
+\begin{align}
+  \sum_{k,n,n'} b_n b_{n'} e^{2\pi ik\p{n-n'}/N}
+  = \sum_{n,n'} b_n b_{n'} \sum_{k=1}^{N-1} e^{2\pi ik\p{n-n'}/N}
+  = \sum_{n,n'} b_n b_{n'} \sum_{k=0}^{N-1} e^{2\pi ik\p{n-n'}/N},
+\end{align}
+where the sum over $k$ vanishes for $n\ne n'$ and equals $N$ when
+$n=n'$, so
+\begin{align}
+  \sum_{k,n,n'} b_n b_{n'} e^{2\pi ik\p{n-n'}/N}
+  = N \sum_n b_n^2 = N^2 \tilde h^2,
   &&
-  \chi \equiv \f{\tilde h^2}{N\p{N-1}\xi},
+  \tilde h^2 \equiv \f1N \sum_n b_n^2 = \f1N \sum_n \p{h_n - \bar h}^2.
+  \label{eq:sum_knn}
+\end{align}
+We therefore have that
+\begin{align}
+  H_{\t{free,eff}}^{(2)}
+  = -\sum_m \f{\p{N/2}^2-m^2}{\p{N-1}\eta U}~ \tilde h^2 \op{m},
+\end{align}
+where the term $\p{N/2}^2$ term contributes a global energy shift
+which we can neglect, while the $m^2$ term is proportional to
+$m^2\op{m}=S_z^2$.  In total, the effective Hamiltonian is thus
+\begin{align}
+  H_{\t{free,eff}} = - \bar h S_z - \Omega S_x + \chi S_z^2,
   &&
-  \tilde h^2 \equiv \f1N \sum_n b_n^2,
-\end{align}
-and in turn
-\begin{align}
-  H_{\t{eff}} = -\bar h S_z - \Omega S_x + \chi S_z^2.
+  \chi \equiv \f{\tilde h^2}{\p{N-1}\eta U}.
 \end{align}

### 2018-07-19 soc_squeezing/soc_squeezing_notes.tex
made notes cover the continuous-drive protocol before the pulsed one, and added section on benchmarkin the OAT and TAT Hamiltonians against FH simulations

@@
 
-\section{Effective spin squeezing}
+\section{Spin squeezing and one-axis twisting}
 \label{sec:OAT}
@@
 \begin{align}
-  H_{\t{OAT}}^z = \chi S_z^2.
+  H_{\t{OAT}}^z = \chi S_z^2,
   \label{eq:H_OAT}
 \end{align}
-An example set of parameters and values for one-axis twisting via
+which generates spin squeezing dynamics with metrological
+applications.  Specifically, one-axis twisting allows for measurements
+of collective spin in the plane orthogonal to the mean spin vector
+$\bk{\v S}$ with a spin fluctuation noise floor which scales as
+$\sim1/N^{2/3}$ (additional info in Peiru's notes\note{(?)}).
+
+An example set of parameters for realizing one-axis twisting via
 $H_{\t{OAT}}^z$ with strontium-87 in a 1-D magic-wavelength lattice is
 provided in Table \ref{tab:parameters}, along with the optimal
-one-axis twisting squeezing time
+squeezing time
 $t_{\t{opt}}^{\t{OAT}}\sim N^{-2/3}\chi^{-1}\sim\eta
@@
 
-\section{Two-axis twisting: pulsed drive}
+\section{Two-axis twisting: continuous drive}
+\label{sec:continuous_drive}
 
-While one-axis twisting provides a spin fluctuation noise floor which
-scales with atom number as $\sim1/N^{2/3}$, it is possible to beat
-this limit and achieve a noise floor scaling as $\sim1/N$ via
-so-called two-axis twisting.  We can transform the one-axis twisting
-Hamitlonian in \eqref{eq:H_OAT} into an effective two-axis twisting
-Hamiltonian by using a pulsed drive sequence developed in
-ref.~\cite{liu2011spin}.  The necessary drive sequence requires the
-pulsed application of a clock laser, which turns on a collective spin
+We can go beyond the $\sim1/N^{2/3}$ one-axis twisting noise floor and
+achieve an improved $\sim1/N$ noise floor via so-called two-axis
+twisting, which can be achieved with external driving protocols.  By
+turning on a clock laser resonant on the electronic transition of the
+atoms, we can realize the Hamiltonian
+\begin{align}
+  H_{\t{drive}} = -\Omega \sum_n s_n^x + H_{\t{OAT}}^z
+  = -\Omega S_x + \chi S_z^2,
+\end{align}
+where the drive strength $\Omega$ may generally depend on time, but
+must always be much smaller in magnitude than the collective spin gap
+$\eta U$ in order preserve the validity of the effective one-axis
+twisting Hamiltonian (see Appendix \ref{sec:squeezing_derivation}).
+Following the prescription in ref.~\cite{huang2015twoaxis} we now
+modulate the drive as $\Omega\p{t}=\beta\omega\cos\p{\omega t}$ with
+$\omega\gg N\chi$, and move into the rotating frame of
+$-\Omega\p{t}S_x$.  After a single secular approximation which relies
+only on $\omega\gg N\chi$, this procedure results in the effective
 Hamiltonian
-\begin{align}
-  H_{\t{clock}} = -\Omega \sum_n s_n^x = -\Omega S_x.
-\end{align}
-If $\abs{\Omega}\gg N\chi$, then we can apply $H_{\t{clock}}$ for
-short times $\theta/\abs{\Omega}\ll\p{N\chi}^{-1}$ (during which we
-can neglect $H_{\t{eff}}$) to realize collective rotations of the form
-$\exp\p{\pm i\theta S_x}$.  Thus
-\begin{enumerate*}[label=(\roman*)]
-\item acting with $\exp\sp{i\p{\pi/2}S_x}$,
-\item waiting for a time $2\tau$,
-\item acting with $\exp\sp{-i\p{\pi/2}S_x}$, and
-\item waiting for a time $\tau$
-\end{enumerate*}
-realizes the unitary
-\begin{align}
-  U_1 = \exp\p{-i\tau\chi S_z^2} \exp\p{-i\f{\pi}{2}S_x}
-  \exp\p{-i2\tau\chi S_z^2} \exp\p{i\f{\pi}{2}S_x}
-  = \exp\p{-i\tau\chi S_z^2} \exp\p{-i2\tau\chi S_y^2}.
-\end{align}
-If $\tau\chi\ll1$, then
-\begin{align}
-  U_1 \approx \exp\sp{-i\tau\chi\p{S_z^2 + 2S_y^2}}
-  = \exp\sp{-i\tau\chi\p{S^2 + S_y^2 - S_x^2}}
-  \simeq \exp\sp{-i\tau\chi\p{S_y^2 - S_x^2}},
-\end{align}
-where $\simeq$ denotes equality up to an overall phase, which holds
-above because $S^2$ is conserved by all dynamics.  Repeating the above
-sequence $n$ times for a time $t=3n\tau$ thus results in the unitary
-\begin{align}
-  U = \exp\sp{-in\tau\chi\p{S_y^2 - S_x^2}}
-  = \exp\sp{-i\p{t/3}\chi\p{S_y^2 - S_x^2}},
-\end{align}
-which is equivalently generated by evolution for a time $t$ under the
-two-axis countertwisting Hamiltonian
-\begin{align}
-  H_{\t{TAT}}^{y,x} = \f{\chi}{3}\p{S_y^2 - S_x^2}.
-  \label{eq:H_TAT_pulse}
-\end{align}
-Despite the factor of 1/3 in the effective two-axis countertwisting
-strength, the optimal squeezing time for this two-axis twisting
-Hamiltonian is generally close to that for one-axis twisting with
-$H_{\t{OAT}}^z$ in \eqref{eq:H_OAT} \cite{liu2011spin}.
-
-\note{[MP: In practice, of course, the gates $\exp\p{i\theta S_x}$
-  take a nonzero amount of time to implement.  For the purposes of
-  simulation, we can probably approximate these gates by square pulses
-  centered on the appropriate time.  I can do an error analysis on
-  this sequence to determine the effect of the $\O\p{\tau^2\chi^2}$
-  terms in the exponent of $U_1$, but it is probably simpler to just
-  try this out in simulations.]}
-
-\note{[MP: I expect this pulsed-drive protocol to be worse than the
-  continuous-drive protocol below.  If we decide that we are not
-  interested in the pulsed drive, we can merely mention it and throw
-  in a citation of ref.~\cite{liu2011spin} without devoting an entire
-  section to the subject.]}
-
-\section{Two-axis twisting: continuous drive}
-
-Rather than pulsing the clock laser, we can apply a continuous drive
-to realize the Hamiltonian
-\begin{align}
-  H_{\t{drive}} = H_{\t{free}} + H_{\t{clock}}
-  = -\f{U}{L} \v S\c\v S - \bar h S_z - \Omega S_x - \sum_n b_n s_n^z,
-\end{align}
-where the drive strength $\Omega$ may generally depend on time.  As
-shown in Appendix \ref{sec:effective_squeezing_drive}, the addition of
-a drive with $\abs{\Omega}\ll\eta\abs{U}=N\abs{U}/L$ does not modify
-the squeezing behavior of the spin-orbit coupling terms in
-$H_{\t{drive}}$, whose effective action on the $S=N/2$ manifold is
-captured by
-\begin{align}
-  H_{\t{drive,eff}} = -\bar h S_z - \Omega S_x + \chi S_z^2.
-  \label{eq:H_drive_eff}
-\end{align}
-To follow the continuous-drive two-axis twisting prescription in
-ref.~\cite{huang2015twoaxis}, we now
-\begin{enumerate*}[label=(\roman*)]
-\item blue-detune the clock laser by $\bar h$ and shift into an
-  appropriate rotating frame to eliminate the $\sim S_z$ term in
-  \eqref{eq:H_drive_eff},
-\item modulate the drive as $\Omega\p{t}=\beta\omega\cos\p{\omega t}$
-  with $\omega\gg N\chi$, and
-\item move into the rotating frame of $-\Omega\p{t}S_x$.
-\end{enumerate*}
-After a secular approximation which relies on $\omega\gg N\chi$, this
-procedure results in the effective Hamiltonian
 \begin{align}
@@
 \end{align}
-where $\simeq$ denotes equality up to a global energy shift. Given the
-set of parameters in Table \ref{tab:parameters}, the two-axis twisting
-Hamiltonian $H_{\t{TAT}}^{z,x}$ can be achieved with a drive frequency
-$\omega=N\abs{\chi U/L}^{1/2}\approx27\times2\pi$ Hz and modulation
-index $\beta=0.906$.  The optimal squeezing time for this protocol is
-$t_{\t{opt}}^{\t{TAT}}\approx0.82$ seconds.  A comparison of squeezing
-through $H_{\t{OAT}}^z$ and $H_{\t{TAT}}^{z,x}$ for the initial state
-$\ket{Y}\equiv\bigotimes_n\p{\ket{\g}_n+i\ket{\e}_n}/\sqrt2$ via the
-parameters in Table \ref{tab:parameters} is provided in Figure
-\ref{fig:squeezing_comparison}.  The figure of merit for squeezing is
-the normalized minimal variance of spin in the plane orthogonal to the
-mean spin vector $\bk{\v S}$:
+where $\simeq$ denotes equality up to a global energy shift.  Given
+the set of parameters in Table \ref{tab:parameters}, the two-axis
+twisting Hamiltonian $H_{\t{TAT}}^{z,x}$ can be achieved with a drive
+frequency $\omega=\abs{\eta U\chi}^{1/2}\approx27\times2\pi$ Hz and
+modulation index $\beta\approx0.906$, and has an optimal squeezing
+time of $t_{\t{opt}}^{\t{TAT}}\approx0.82$ seconds.  The figure of
+merit for squeezing is the normalized minimal variance of spin in the
+plane orthogonal to the mean spin vector $\bk{\v S}$:
 \begin{align}
@@
 \end{align}
-for $\abs{\hat{\v n}}=1$.
+for $\abs{\hat{\v n}}=1$.  A comparison of this squeezing parameter
+through $H_{\t{OAT}}^z$ and $H_{\t{TAT}}^{z,x}$ for the initial state
+$\ket{Y}\equiv\bigotimes_n\p{\ket{\g}_n+i\ket{\e}_n}/\sqrt2$ via the
+parameters in Table \ref{tab:parameters} is provided in Figure
+\ref{fig:squeezing_comparison_TAT}.
 
@@
     with $\omega=27\times2\pi$ Hz and $\beta=0.906$.}
-  \label{fig:squeezing_comparison}
+  \label{fig:squeezing_comparison_TAT}
 \end{figure}
 
-\section{Discussions}
+\section{Two-axis twisting: pulsed drive}
 
-\note{[MP: We should clarify in the notes what exactly the conditions
-  are for the effective Hamiltonians we derive, i.e.~the spin model
-  and the one-axis twisting Hamiltonian.]}
+Two-axis twisting can also be achieved using the pulsed-drive sequence
+described in ref.~\cite{liu2011spin}.  This sequence requires the
+drive strength $\Omega$ much greater than the collective spin gap, or
+$\abs{\Omega}\gg\eta\abs{U}$, which allows us to generate collective
+spin rotations of the form $\exp\p{\pm i\theta S_x}$ with short pulses
+during which we can neglect any free evolution of the atoms.  Thus
+\begin{enumerate*}[label=(\roman*)]
+\item acting with $\exp\sp{i\p{\pi/2}S_x}$,
+\item waiting for a time $2\tau/3$,
+\item acting with $\exp\sp{-i\p{\pi/2}S_x}$, and
+\item waiting for a time $\tau/3$
+\end{enumerate*}
+realizes the unitary
+\begin{align}
+  U_\tau
+  &= \exp\p{-i\f13\tau\chi S_z^2} \exp\p{-i\f{\pi}{2}S_x}
+  \exp\p{-i\f23\tau\chi S_z^2} \exp\p{i\f{\pi}{2}S_x} \\
+  &= \exp\p{-i\f13\tau\chi S_z^2} \exp\p{-i\f23\tau\chi S_y^2},
+\end{align}
+where if $\tau\chi\ll1$, then
+\begin{align}
+  U_\tau \approx \exp\sp{-i\f13\tau\chi\p{S_z^2 + 2S_y^2}}
+  = \exp\p{-i\tau H_{\t{TAT}}^{y,x}}.
+\end{align}
+Such a sequence thus generates evolution which is equivalent to simply
+evolving for a time $\tau$ under the two-axis twisting Hamiltonian
+$H_{\t{TAT}}^{y,x}$.
 
-\note{[MP] Future ideas:} Make use of a cavity to enhance the one-axis
-twisting strength \cite{hu2017vacuum}?
+\section{Benchmarking}
 
+For small system sizes, we can verify the validity of the effective
+one- and two-axis twisting Hamiltonians via direct simulations of a
+Fermi-Hubbard (FH) model, using the spin-orbit coupling Hamiltonian
+$H_{\t{SOC}}$ and interaction Hamiltonian $H_{\t{int}}$ respectively
+given in \eqref{eq:H_SOC_start} and \eqref{eq:H_int_full}.  Figure
+\ref{fig:squeezing_comparison_FH} shows a comparison of the squeezing
+parameter $\xi^2$ after evolution of the initial state $\ket{Y}$ for
+$N=8$ particles in a 1-D lattice with $L=8$ sites (and all remaining
+parameters as given in Table \ref{tab:parameters}) after evolution
+under
+\begin{enumerate*}[label=(\roman*)]
+\item the one-axis twisting Hamiltonian $H_{\t{OAT}}^z$,
+\item the two-axis twisting Hamiltonian $H_{\t{OAT}}^{z,x}$,
+\item free evolution of the Fermi-Hubbard model under
+  $H_{\t{SOC}}+H_{\t{int}}$, and
+\item driven evolution of the Fermi-Hubbard model using the
+  continuous-drive protocol in Section \ref{sec:continuous_drive} with
+  drive frequency $\omega=\abs{\eta U\chi}^{1/2}$ and modulation index
+  $\beta=0.906$.
+\end{enumerate*}
+These simulation results show essentially perfect agreement between
+the Fermi-Hubbard model and the one- and two-axis twisting
+Hamiltonians.
+
+\begin{figure}
+  \centering \includegraphics{squeezing_comparison_FH.pdf}
+  \caption{Benchmarking squeezing under the effective one- and
+    two-axis Hamiltonians $H_{\t{OAT}}^z$ and $H_{\t{TAT}}^{z,x}$
+    against direct simulations of the Fermi-Hubbard model with
+    $N=L=8$.}
+  \label{fig:squeezing_comparison_FH}
+\end{figure}
+
+% \note{[MP] Future ideas:} Make use of a cavity to enhance the
+% one-axis twisting strength \cite{hu2017vacuum}?
+
+\newpage
 \appendix

### 2018-07-19 soc_squeezing/soc_squeezing_notes.tex
added example figure for squeezing in 2D

@@
     Transverse lattice depths & $V_T$ & 60 $E_R$ \\
+    SOC strength & $\phi$ & $\pi/50$ \\
     Lattice sites & $L$ & 100 \\
-    Atom number & $N$ & 50 \\
-    SOC strength & $\phi$ & $\pi/50$ \\ \hline\hline
+    Atom number & $N$ & 50 \\ \hline\hline
     Tunneling rate & $J_0$ & $\approx300\times2\pi$ Hz \\
@@
 one- and two-axis twisting Hamiltonians via direct simulations of a
-Fermi-Hubbard (FH) model, using the spin-orbit coupling Hamiltonian
-$H_{\t{SOC}}$ and interaction Hamiltonian $H_{\t{int}}$ respectively
-given in \eqref{eq:H_SOC_start} and \eqref{eq:H_int_full}.  Figure
-\ref{fig:squeezing_comparison_FH} shows a comparison of the squeezing
-parameter $\xi^2$ after evolution of the initial state $\ket{Y}$ for
-$N=8$ particles in a 1-D lattice with $L=8$ sites (and all remaining
-parameters as given in Table \ref{tab:parameters}) after evolution
-under
+single-band Fermi-Hubbard (FH) model, using the spin-orbit coupling
+Hamiltonian $H_{\t{SOC}}$ and interaction Hamiltonian $H_{\t{int}}$
+respectively given in \eqref{eq:H_SOC_start} and
+\eqref{eq:H_int_full}.  Figure \ref{fig:squeezing_comparison_FH} shows
+a comparison of the squeezing parameter $\xi^2$ after evolution of the
+initial state $\ket{Y}$ for $N=8$ particles in a 1-D lattice with
+$L=8$ sites (and all remaining parameters as given in Table
+\ref{tab:parameters}) after evolution under
 \begin{enumerate*}[label=(\roman*)]
@@
 the Fermi-Hubbard model and the one- and two-axis twisting
-Hamiltonians.
+Hamiltonians for the given parameters.
 
 \begin{figure}
-  \centering \includegraphics{squeezing_comparison_FH.pdf}
-  \caption{Benchmarking squeezing under the effective one- and
-    two-axis Hamiltonians $H_{\t{OAT}}^z$ and $H_{\t{TAT}}^{z,x}$
-    against direct simulations of the Fermi-Hubbard model with
-    $N=L=8$.}
+  \centering
+  \includegraphics{squeezing_comparison_FH.pdf}
+  \caption{Benchmarking squeezing under the one- and two-axis
+    Hamiltonians $H_{\t{OAT}}^z$ and $H_{\t{TAT}}^{z,x}$ against
+    direct simulations of the Fermi-Hubbard model with $N=L=8$ and the
+    remaining lattice/SOC parameters as given in Table
+    \ref{tab:parameters}.}
   \label{fig:squeezing_comparison_FH}
@@
 
-% \note{[MP] Future ideas:} Make use of a cavity to enhance the
-% one-axis twisting strength \cite{hu2017vacuum}?
+\section{Last thoughts}
+
+It may be possible to further enhance the twisting strength $\chi$ by
+placing the atoms in an optical cavity, as in
+ref.~\cite{hu2017vacuum}.  Both squeezing times and squeezing
+strengths might also be further enhanced by using a 2-D lattice,
+although more benchmarking is necessary to verify the validity of the
+collective spin model in 2-D.  As shown in Figure
+\ref{fig:squeezing_comparison_2D},
+
+\begin{figure}
+  \centering
+  \includegraphics{squeezing_comparison_2D.pdf}
+  \caption{Squeezing via one- and two-axis twisting in a 2-D lattice
+    with $100\times100$ sites at half filling, with the remaining
+    lattice/SOC parameters given in Table \ref{tab:parameters}.}
+  \label{fig:squeezing_comparison_2D}
+\end{figure}
 

### 2018-07-20 soc_squeezing/soc_squeezing_notes.tex
removed incorrect prefactor on Dicke states; use |m> \propto ... instead
modified derivation of S \dot S Hamiltonian added capability to create a fermionic operator by index added method creating projector onto states with an occupied lattice site changed order of arguments for Hamiltonians in FH methods

@@
   \sp{\p{\f{\1_q-\sigma_q^z}{2}} \p{\f{\1_p+\sigma_p^z}{2}}
-    - \sigma_q^+ \sigma_p^-}.
+    - \sigma_q^+ \sigma_p^-}
+  = \f{U}{L} \sum_{p,q}\f14\p{\1_q \1_p - \v\sigma_q\c\v\sigma_p},
+  \label{eq:H_int_pauli}
 \end{align}
-Up to a global shift in energy, this result is equivalently
+with
 \begin{align}
-  H_{\t{int}}
-  = - \f14 \f{U}{L} \sum_{p,q} \v{\sigma}_q\c\v{\sigma}_p
-  = - \f{U}{L} \sum_{p,q} \v s_q \c\v s_p
-  = - \f{U}{L} \v S \c \v S,
+  \v{\sigma}_p \equiv \p{\sigma_p^x,\sigma_p^y,\sigma_p^z}.
 \end{align}
-where
+For a fixed total particle number $N$, the identity operators in
+\eqref{eq:H_int_pauli} contribute only a global shift in energy, which
+allows us to more simply write
 \begin{align}
-  \v{\sigma}_p
-  \equiv \p{\sigma_p^x,\sigma_p^y,\sigma_p^z},
-  &&
-  \v s_p \equiv \p{s_p^x,s_P^y,s_p^z},
+  H_{\t{int}} = - \f{U}{L} \v S\c\v S,
   &&
   \v S \equiv \f12 \sum_p \v{\sigma}_p.
-  \label{eq:H_int}
 \end{align}
-For a fixed particle number $N$, the collective spin Hamiltonian
-$H_{\t{int}}$ has electronic eigenstates $\set{\ket{Sm}}$ with total
-(pseudo-)spin $S$ and projection $m$ onto a quantization axis; the
-corresponding energies are $\bk{Sm|H_{\t{int}}|S,m}=-\p{U/L}S\p{S+1}$.
-In particular, the ground-state $S=N/2$ manifold is spanned by the
-Dicke states
+This Hamiltonian has electronic eigenstates $\set{\ket{Sm}}$ with
+total (pseudo-)spin $S$ and projection $m$ onto a quantization axis;
+the corresponding energies are
+$\bk{Sm|H_{\t{int}}|S,m}=-\p{U/L}S\p{S+1}$.  In particular, the
+ground-state $S=N/2$ manifold is spanned by the Dicke states
 \begin{align}
-  \ket{m} \equiv \ket{N/2,m}
-  = {N \choose N/2+m}^{-1/2} S_+^{N/2+m} \ket{\g}^{\otimes N},
+  \ket{m} \equiv \ket{N/2,m} \propto S_+^{N/2+m} \ket{\g}^{\otimes N},
   &&

### 2018-07-20 soc_squeezing/scripts/fermi_hubbard_methods.py

@@ -42,8 +42,8 @@ def spacial_basis(L):
 # iterable for basis of fock states; returns index of fock state
 #   and a tuple specifying the occupied single-particle states (by index)
 def fock_state_basis(L, N):
-    fock_states = int(binomial(2*product(L),N))
-    return zip(range(fock_states), itertools.combinations(range(2*product(L)),N))
+    hilbert_dim = int(binomial(2*product(L),N))
+    return zip(range(hilbert_dim), itertools.combinations(range(2*product(L)),N))
 
 # get operator representation of annihilation operators in the fock basis
 # depth is the number of times we need to apply annihilation operators
@@ -150,12 +150,26 @@ class c_op:
     # index of the state addressed by this operator
     def index(self, L):
         L = np.array(L, ndmin = 1)
-        site_index = sum([ self.q[ii] * product([ L[jj] for jj in range(ii+1,len(L)) ])
+        site_index = sum([ ( self.q[ii] % L[ii] )
+                           * product([ L[jj] for jj in range(ii+1,len(L)) ])
                            for ii in range(len(self.q)) ])
         return 2 * site_index + self.spin_up
 
     def vector(self, L, N):
         return c_seq(self).vector(L, N)
+
+# return a fermionic operator by the value of its index in a given lattice geometry
+def c_op_idx(index, L):
+    L = np.array(L, ndmin = 1)
+    dim = len(L)
+    spin_up = index % 2
+    q = np.zeros(dim)
+    for ii in range(dim-1,-1,-1):
+        q_gtr_cont = sum([ q[jj] * product([ L[kk] for kk in range(jj+1,dim) ])
+                           for jj in range(ii+1,dim) ])
+        q[ii] = ( ( index // 2 - q_gtr_cont )
+                  // product([ L[jj] for jj in range(ii+1,dim) ]) ) % L[ii]
+    return c_op(q,spin_up)
 
 # product of creation / annihilation operators
 class c_seq:
@@ -213,6 +227,8 @@ class c_seq:
             sys.exit("error: adding invalid type to c_seq:", type(other))
 
     def __sub__(self, other): return self + (-1) * other
+
+    def __neg__(self): return -1 * self
 
     # sort operators change sign of prefactor if appropriate
     def sort(self):
[…]

### 2018-07-24 olc_soc/soc_interactions.tex
removed unnecessary text and fixed 'extensive with N' to 'extensive with L'

@@
   we use driving schemes to match the mean band energies, thereby
-  effective setting $\bar E_1=\bar E_0$; these terms can be neglected
-  in the absence of such driving.
-\item The number of nonvanishing terms in $\v S_n$ is proportional to
-  the occupation of band $n$.  If there is a large population of atoms
-  in the ground band and a small population of atoms in the excited
-  band, then we can safely neglect the $\sim\v S_1\c\v S_1$ term,
-  which is doubly small compared to e.g. $\v S_0\c\v S_1$.
-\item Similarly, if the ground band is nearly filled with atoms in one
-  clock state and the excited band only sparsely populated with atoms
-  in the opposite clock state, then we can neglect the terms
-  $\sim b_{-q,\bar\sigma}^\dag b_{q\bar\sigma} a_{p+2q,\sigma}^\dag
-  a_{p\sigma}$, as $b_{-q,\bar\sigma}^\dag b_{q\bar\sigma}$ manifestly
-  addresses a small number of atoms while
-  $a_{p+2q,\sigma}^\dag a_{p\sigma}$ is only nonzero for quasi-momenta
-  $p$ with $p+2q$ unoccupied.
+  effectively setting $\bar E_1=\bar E_0$; these terms can be
+  neglected in the absence of such driving.
 \item If there is a nonzero population of atoms in the excited band,
@@
   excitations across all quasi-momenta (on a time scale
-  $\sim1/G_{0,1}$ which is extensive with $N$).
+  $\sim1/G_{0,1}$ which is extensive with $L$).
 \end{enumerate}

### 2018-07-26 soc_squeezing/soc_squeezing_notes.tex
modified some figure parameters and added 'bare' lattice Hamiltonian prior to Hubbard model approximation

@@
   H_{\t{SOC}}
-  = -2J_0\sum_{q,s} \cos\p{q+s\phi/2} \hat c_{qs}^\dag \hat c_{qs}
+  = \sum_{q,s} E_{qs\phi} \hat c_{qs}^\dag \hat c_{qs}
+  \approx -2 J_0\sum_{q,s} \cos\p{q+s\phi/2}
+  \hat c_{qs}^\dag \hat c_{qs}
   \label{eq:H_SOC_start}
@@
   G \equiv \f{4\pi a_{\e\g^-}}{m_A},
+  \label{eq:H_int_full}
 \end{align}
@@
 \end{align}
-which motivates the definitions
+which motivates the definition
 \begin{align}
@@
   \hat c_{q,\g} \hat c_{p,\e}.
-  \label{eq:H_int_full}
+  \label{eq:H_int_uniform}
 \end{align}
@@
 \end{enumerate*}
-The surviving terms in \eqref{eq:H_int_full} are thus
+The surviving terms in \eqref{eq:H_int_uniform} are thus
 \begin{align}
@@
     Parameter & Symbol & Value \\ \hline\hline
-    Primary lattice depth & $V_0$ & 4 $E_R$ \\
-    Transverse lattice depths & $V_T$ & 60 $E_R$ \\
+    Primary lattice depth & $V_0$ & 5 $E_R$ \\
+    Transverse lattice depths & $V_T$ & 50 $E_R$ \\
     SOC strength & $\phi$ & $\pi/50$ \\
     Lattice sites & $L$ & 100 \\
-    Atom number & $N$ & 50 \\ \hline\hline
-    Tunneling rate & $J_0$ & $\approx300\times2\pi$ Hz \\
-    On-site interaction strength & $U$ & $\approx1.4\times2\pi$ kHz \\
-    Squeezing strength & $\chi$ & $\approx21\times2\pi$ mHz \\
-    Optimal squeezing time & $t_{\t{opt}}^{\t{OAT}}$
-    & $\approx0.59$ seconds \\ \hline
+    Atom number & $N$ & 100 \\ \hline\hline
+    Tunneling rate & $J_0$ & $\approx230\times2\pi$ Hz \\
+    On-site interaction strength & $U$ & $\approx1.3\times2\pi$ kHz \\
+    Twisting strength & $\chi$ & $\approx3.2\times2\pi$ mHz \\
+    Optimal OAT squeezing time & $t_{\t{opt}}^{\t{OAT}}$
+    & $\approx2.5$ seconds \\ \hline
   \end{tabular}
@@
 twisting Hamiltonian $H_{\t{TAT}}^{z,x}$ can be achieved with a drive
-frequency $\omega=\abs{\eta U\chi}^{1/2}\approx27\times2\pi$ Hz and
+frequency $\omega=\abs{\eta U\chi}^{1/2}\approx20\times2\pi$ Hz and
 modulation index $\beta\approx0.906$, and has an optimal squeezing
-time of $t_{\t{opt}}^{\t{TAT}}\approx0.82$ seconds.  The figure of
+time of $t_{\t{opt}}^{\t{TAT}}\approx3.2$ seconds.  The figure of
 merit for squeezing is the normalized minimal variance of spin in the
@@
     \ref{tab:parameters} and the continuous two-axis twisting protocol
-    with $\omega=27\times2\pi$ Hz and $\beta=0.906$.}
+    with $\omega\approx\abs{\eta U\chi}^{1/2}$ and $\beta=0.906$.}
   \label{fig:squeezing_comparison_TAT}
@@
 initial state $\ket{Y}$ for $N=8$ particles in a 1-D lattice with
-$L=8$ sites (and all remaining parameters as given in Table
-\ref{tab:parameters}) after evolution under
+$L=8$ sites (and remaining system parameters as given in the top half
+of Table \ref{tab:parameters}) after evolution under
 \begin{enumerate*}[label=(\roman*)]
@@
     direct simulations of the Fermi-Hubbard model with $N=L=8$ and the
-    remaining lattice/SOC parameters as given in Table
+    remaining system parameters as given in the top half of Table
     \ref{tab:parameters}.}
@@
 placing the atoms in an optical cavity, as in
-ref.~\cite{hu2017vacuum}.  Both squeezing times and squeezing
-strengths might also be further enhanced by using a 2-D lattice,
-although more benchmarking is necessary to verify the validity of the
-collective spin model in 2-D.  As shown in Figure
+ref.~\cite{hu2017vacuum}.  Both squeezing times and twisting strengths
+might also be further enhanced by using a 2-D lattice, although more
+benchmarking is necessary to verify the validity of the collective
+spin model in 2-D.  As shown in Figure
 \ref{fig:squeezing_comparison_2D},
@@
   \caption{Squeezing via one- and two-axis twisting in a 2-D lattice
-    with $100\times100$ sites at half filling, with the remaining
+    with $100\times100$ sites at unit filling, with the remaining
     lattice/SOC parameters given in Table \ref{tab:parameters}.}

### 2018-07-28 soc_squeezing/soc_squeezing_notes.tex
more parameter tweaking for the squeezing protocols, but this time using the depth exclusion plots

@@
     Primary lattice depth & $V_0$ & 5 $E_R$ \\
-    Transverse lattice depths & $V_T$ & 50 $E_R$ \\
-    SOC strength & $\phi$ & $\pi/50$ \\
+    Transverse lattice depths & $V_T$ & 60 $E_R$ \\
+    SOC strength & $\phi$ & $\pi/25$ \\
     Lattice sites & $L$ & 100 \\
@@
     Tunneling rate & $J_0$ & $\approx230\times2\pi$ Hz \\
-    On-site interaction strength & $U$ & $\approx1.3\times2\pi$ kHz \\
-    Twisting strength & $\chi$ & $\approx3.2\times2\pi$ mHz \\
+    On-site interaction strength & $U$ & $\approx1.5\times2\pi$ kHz \\
+    Twisting strength & $\chi$ & $\approx11\times2\pi$ mHz \\
     Optimal OAT squeezing time & $t_{\t{opt}}^{\t{OAT}}$
-    & $\approx2.5$ seconds \\ \hline
+    & $\approx0.70$ seconds \\ \hline
   \end{tabular}
@@
 twisting Hamiltonian $H_{\t{TAT}}^{z,x}$ can be achieved with a drive
-frequency $\omega=\abs{\eta U\chi}^{1/2}\approx20\times2\pi$ Hz and
+frequency $\omega=\abs{\eta U\chi}^{1/2}\approx41\times2\pi$ Hz and
 modulation index $\beta\approx0.906$, and has an optimal squeezing
-time of $t_{\t{opt}}^{\t{TAT}}\approx3.2$ seconds.  The figure of
+time of $t_{\t{opt}}^{\t{TAT}}\approx0.89$ seconds.  The figure of
 merit for squeezing is the normalized minimal variance of spin in the
@@
 benchmarking is necessary to verify the validity of the collective
-spin model in 2-D.  As shown in Figure
-\ref{fig:squeezing_comparison_2D},
+spin model in 2-D.  Figure \ref{fig:squeezing_comparison_2D} shows a
+comparison of squeezing strengths via one- and two-axis squeezing in a
+2-D lattice with $100\times100$ sites.
 
@@
   \includegraphics{squeezing_comparison_2D.pdf}
-  \caption{Squeezing via one- and two-axis twisting in a 2-D lattice
-    with $100\times100$ sites at unit filling, with the remaining
-    lattice/SOC parameters given in Table \ref{tab:parameters}.}
+  \caption{Squeezing via one- and two-axis twisting in a $V_0=7E_R$
+    depth 2-D lattice with $100\times100$ sites at unit filling
+    ($N=10^4$), transverse confinement depth $V_T=80E_R$, and
+    spin-orbit coupling strength $\phi=\pi/25$.}
   \label{fig:squeezing_comparison_2D}

### 2018-08-15 soc_squeezing/cumulant_expansions.tex
added cumulant expansion notes

\documentclass[aps,notitlepage,nofootinbib,11pt]{revtex4-1}

% linking references
\usepackage{hyperref}
\hypersetup{
  breaklinks=true,
  colorlinks=true,
  linkcolor=blue,
  filecolor=magenta,
  urlcolor=cyan,
}

%%% symbols, notations, etc.
\usepackage{physics,braket,bm,commath,amssymb} % physics and math
\renewcommand{\t}{\text} % text in math mode
\newcommand{\f}[2]{\dfrac{#1}{#2}} % shorthand for fractions
\newcommand{\p}[1]{\left(#1\right)} % parenthesis
\renewcommand{\sp}[1]{\left[#1\right]} % square parenthesis
\renewcommand{\set}[1]{\left\{#1\right\}} % curly parenthesis
\renewcommand{\v}{\bm} % bold vectors
\newcommand{\uv}[1]{\v{\hat{#1}}} % unit vectors
\renewcommand{\c}{\cdot} % inner product
\newcommand{\bk}{\Braket} % shorthand for braket notation
\renewcommand{\d}{\text{d}} % "d" for integration measure

\newcommand{\D}{\mathcal{D}}
\renewcommand{\O}{\mathcal{O}}

\usepackage{dsfont} % for identity operator
\newcommand{\1}{\mathds{1}}

\usepackage[inline]{enumitem} % for inline enumeration

%%% figures
\usepackage{graphicx} % for figures
\usepackage{grffile} % help latex properly identify figure extensions
\usepackage[caption=false]{subfig} % subfigures (via \subfloat[]{})
\graphicspath{{./figures/}} % set path for all figures

% for strikeout text
% normalem included to prevent underlining titles in the bibliography
\usepackage[normalem]{ulem}

% for leaving notes in the text
\newcommand{\note}[1]{\textcolor{red}{#1}}

\begin{document}

\title{Solving Collective Spin Hamiltonians with Decoherence}

\author{Michael A. Perlin}

\maketitle

Previous work has shown how to realize various spin-squeezing
Hamiltonians with ultracold atomic systems.  While it is
straightforward to simulate the dynamics generated by these
Hamiltonians in the context of a closed quantum system, here we try to
develop an understanding of their the squeezing behavior in the
presence of decoherence.  In particular, we will be considering
dynamics induced by the collective one-axis twisting (OAT), two-axis
twisting (TAT), and transverse field (TVF) Hamiltonians of the form
\begin{align}
  H_{\t{OAT}} = \chi S_z^2,
  &&
  H_{\t{TAT}} = \chi \p{S_z^2 - S_y^2},
  &&
  H_{\t{TVF}} = \chi S_z^2 - \Omega S_x,
\end{align}
for spins which decay in an uncorrelated manner at a rate $\gamma$.
The evolution of any Heisenburg operator $\O$ with a Hamiltonian $H$
is thus given by
\begin{align}
  \f{d}{dt} \O
  = i\sp{H,\O}_- + \gamma\D\p{\O},
  &&
  \D\p{\O} \equiv \sum_j\p{\sigma_j^+\O\sigma_j^-
    - \f12\sp{\sigma_j^+\sigma_j^-,\O}_+},
  \label{eq:EOM}
\end{align}
where $\sp{A,B}_\pm\equiv AB\pm BA$ and $j$ indexes an individual
spin.  Our measure of squeezing is
\begin{align}
  \xi^2 \equiv \f{N}{\abs{\bk{\v S}}^2}
  \min_{\uv n\perp\bk{\v S}} \bk{\p{\v S\c\uv n}^2},
  \label{eq:squeezing}
\end{align}
where $N$ is the total number of spins, $\v S$ is the collective spin
vector, and the minimization is performed over all unit vectors
$\uv n$ in the plane orthogonal to the mean spin vector $\bk{\v S}$.
This squeezing parameter is entirely determined by components
$\bk{S_\alpha}$ of the mean spin vector in addition to collective
spin-spin correlators of the form $\bk{S_\alpha S_\beta}$; our task is
therefore to compute these expectation values in all scenarios of
interest.

\section{One-axis twisting}

The case of the one-axis twisting is simplest to consider, as we can
write the Hamiltonian in the form
\begin{align}
  H_{\t{OAT}} = \f14 \chi \sum_{j,k} \sigma_j^z \sigma_k^z
  = \f12 \chi \sum_{j<k} \sigma_j^z \sigma_k^z + \f14 N \chi,
\end{align}
which admits an exact solution previously worked out by Foss-Feig et
al.~in ref.~\cite{foss-feig2013nonequilibrium}.  Adapting exact
expectation values with respect to the initial state $\ket{X}$ which
has all spins pointing in $x$, in terms of $\mu,\nu\in\set{+1,-1}$ we
find that
\begin{align}
  \bk{S_+}
  &= \sum_j\bk{\sigma_j^+}
  = \f{N}{2} e^{-\gamma t/2} \Phi\p{\chi,t}^{N-1}, \label{eq:S+_OAT} \\
  \bk{S_\mu S_z}
  &= \f12\sum_j\bk{\sigma_j^\mu\sigma_j^z}
  + \f12\sum_{j\ne k} \bk{\sigma_j^\mu \sigma_k^z} \\
  &= -\f{\mu}{2}\bk{S_\mu} + \f14 N \p{N-1} e^{-\gamma t/2}
  \Psi\p{\mu\chi,t} \Phi\p{\chi,t}^{N-2}, \\
  \bk{S_\mu S_\nu}
  &= \sum_j \bk{\sigma_j^\mu \sigma_j^\nu}
  + \sum_{j\ne k} \bk{\sigma_j^\mu \sigma_k^\nu} \\
  &= \delta_{\mu,-\nu} \p{\f{N}{2} + \mu \bk{S_z}}
  + \f14 N \p{N-1} e^{-\gamma t}
  \Phi\p{\sp{\mu+\nu}\chi,t}^{N-2}, \label{eq:SS+-_OAT}
\end{align}
where
\begin{align}
  \Phi\p{\zeta,t} \equiv e^{-\gamma t/2}
  \sp{\cos\p{s_\zeta t}+\f12\gamma t~\t{sinc}\p{s_\zeta t}},
  &&
  \Psi\p{\zeta,t} \equiv e^{-\gamma t/2}
  \p{is_\zeta-\gamma/2}t~ \t{sinc}\p{s_\zeta t},
\end{align}
and $s_\zeta \equiv \zeta + i2\gamma$.  The last ingredients we need
to compute spin squeezing at any time are $\bk{S_z}$ and $\bk{S_z^2}$.
As these operators commute with the OAT Hamiltonian, their evolution
is governed entirely by decoherence, such that
\begin{align}
  \f{d}{dt}\p{S_z^n}
  = \gamma \D\p{S_z^n}
  = \gamma \p{S + S_z} \sum_{k=0}^{n-1}
  { n \choose k } \p{-1}^{n-k} S_z^k,
\end{align}
where $S\equiv N/2$ and $\D\p{S_z^n}$ is worked out in Appendix
\ref{sec:decoherence}.  In particular,
\begin{align}
  \f{d}{dt} S_z = -\gamma \p{S + S_z},
  &&
  \f{d}{dt}\p{S_z^2} = \gamma \p{S + S_z} \p{1 - 2S_z}.
\end{align}
The initial conditions $\bk{S_z}_{t=0}=0$ and $\bk{S_z^2}_{t=0}=S/2$
then imply
\begin{align}
  \bk{S_z} = S\p{e^{-\gamma t}-1},
  &&
  \bk{S_z^2} = S \p{1 - \f12 e^{-\gamma t}} e^{-\gamma t} + \bk{S_z}^2.
  \label{eq:Sz_OAT}
\end{align}
The expectation values in \eqref{eq:S+_OAT}-\eqref{eq:SS+-_OAT} and
\eqref{eq:Sz_OAT} are sufficient to compute the squeezing parameter
$\xi^2$ in \eqref{eq:squeezing} at any time throughout evolution of
the initial state $\ket{X}$ under $H_{\t{OAT}}$.

\subsection{Cumulant expansion}

Unlike the one-axis twisting Hamiltonian, the two-axis twisting and
transverse-field Hamiltonians do not admit exact analytical
solutions. We will therefore solve these Hamiltonians approximately
using cumulant expansions.  The basic idea is to find equations of
motion for operators of interest, and at some point truncate
high-order terms to arrive at a closed, reasonably-sized system of
equations.  To benchmark this method, we will test it against exact
one-axis twisting results.

In order to find equations of motion for $S_\mu$, $S_\mu S_z$, and
$S_\mu S_\nu$ (i.e.~where $\mu,\nu\in\set{+1,-1}$), we first compute
the commutator
\begin{align}
  \sp{S_z^2, S_+^\ell S_-^m S_z^n}_-
  = \p{S_z^2 S_+^\ell S_-^m - S_+^\ell S_-^m S_z^2} S_z^n,
\end{align}
where we can use the commutation relation
$\sp{S_z,S_\mu^k}_-=\mu k S_\mu$ (see Appendix \ref{sec:identities})
to push spin-$z$ operators (i.e.~$S_z$) to the right, as
\begin{align}
  S_z S_+^\ell S_-^m
  = S_+^\ell \p{\ell + S_z} S_-^m
  = \ell S_+^\ell S_-^m + S_+^\ell S_z S_-^m
  = S_+^\ell S_-^m \p{\ell - m + S_z},
\end{align}
which implies
\begin{align}
  \sp{S_z^2, S_+^\ell S_-^m S_z^n}_-
  = S_+^\ell S_-^m \p{\sp{\ell - m + S_z}^2 - S_z^2} S_z^n
  = S_+^\ell S_-^m \p{\sp{\ell-m}^2 + 2\sp{\ell-m} S_z} S_z^n.
\end{align}
For the OAT Hamiltonian, we then have that
\begin{align}
  \f{d}{dt} \p{S_+^\ell S_-^m S_z^n}
  = i\chi S_+^\ell S_-^m \p{\sp{\ell-m}^2 + 2\sp{\ell-m} S_z} S_z^n
  + \gamma \D\p{S_+^\ell S_-^m S_z^n},
\end{align}
where $\D\p{S_+^\ell S_-^m S_z^n}$ is worked out in Appendix
\ref{sec:decoherence}.  In order to compute squeezing of the OAT
Hamiltonian in the presence of single-spin decay, we therefore need to
solve the systems of equations defined by
\begin{align}
  \f{d}{dt} \p{S_\mu S_z^n}
  = i\chi S_\mu \p{1 + 2\mu S_z} S_z^n
  + \gamma S_\mu \sp{\p{S - \delta_{\mu,-1} + S_z} \sum_{k=0}^{n-1}
    { n \choose k } \p{-1}^{n-k} S_z^k - \f12 S_z^n}
\end{align}
and
\begin{multline}
  \f{d}{dt} \p{S_\mu S_\nu S_z^n}
  = i\chi S_\mu S_\nu \p{\sp{\mu+\nu}^2 + 2 \sp{\mu+\nu} S_z} S_z^n \\
  + \gamma S_\mu S_\nu
  \sp{\p{S - \delta_{\mu,-1} - \delta_{\nu,-1} + S_z} \sum_{k=0}^{n-1}
    { n \choose k } \p{-1}^{n-k} S_z^k - S_z^n},
\end{multline}
for $\mu\ge\nu$.

\bibliography{\jobname}

\newpage
\appendix

\section{Spin operator identities}
\label{sec:identities}

This text frequently makes use of various spin operator identities; we
collect some particularly useful identities here for reference.
\begin{align}
  \f12 \sp{\sigma_j^z,\sigma_k^\mu}_- = \delta_{jk} \mu \sigma_j^\mu,
  &&
  \sp{S_z,\sigma_j^\mu}_- = \f12 \sp{\sigma_j^z,S_\mu}_-
  = \mu \sigma_j^\mu,
  &&
  \sp{S_z,S_\mu}_- = \mu S_\mu.
  \label{eq:comm_z_mu}
\end{align}
\begin{align}
  \sp{\sigma_j^\mu,\sigma_k^\nu}_-
  = \delta_{jk} \delta_{\mu,-\nu} \mu \sigma_j^z,
  &&
  \sp{S_\mu,\sigma_j^\nu}_- = \sp{\sigma_j^\mu,S_\nu}_-
  = \delta_{\mu,-\nu} \mu \sigma_j^z,
  &&
  \sp{S_+,S_-}_- = 2 S_z.
  \label{eq:comm_mu_nu}
\end{align}
\begin{align}
  S_z S_\mu^n
  = \p{\mu S_\mu + S_\mu S_z} S_\mu^{n-1}
  = \mu S_\mu^n + S_\mu S_z S_\mu^{n-1}
  = 2\mu S_\mu^n + S_\mu^2 S_z S_\mu^{n-2}
  = S_\mu^n \p{\mu n + S_z},
  \label{eq:comm_z_mu_n}
\end{align}
\begin{align}
  \sigma_j^+ \sigma_j^- = \f12\p{\1_j+\sigma_j^z},
  &&
  \sum_j \sigma_j^+ \sigma_j^- = S + S_z.
\end{align}
Here $S\equiv N/2$.

\section{Decoherence of Heisenberg spin operators}
\label{sec:decoherence}

The decoherence operator for uncorrelated single-spin decay at a
uniform rate is
\begin{align}
  \D\p{\O}
  \equiv \sum_j\p{\sigma_j^+\O\sigma_j^-
    - \f12\sp{\sigma_j^+\sigma_j^-,\O}_+}
  = \sum_j \sigma_j^+\O\sigma_j^- - S \O - \f12 \sp{S_z,\O}_+.
\end{align}
We are interested in the effect of this decoherence on the dynamics of
collective spin operators, all of which are determined by operators of
the form $S_+^\ell S_-^m S_z^n$.  We thus expand
\begin{align}
  \D\p{S_+^\ell S_-^m S_z^n}
  = \sum_j \sigma_j^+ S_+^\ell S_-^m S_z^n \sigma_j^-
  - S S_+^\ell S_-^m S_z^n - \f12 \sp{S_z, S_+^\ell S_-^m S_z^n}_+.
  \label{eq:D_z+-}
\end{align}
Considering the first term, we can use a commutation relations from
\eqref{eq:comm_z_mu} to push through and simplify the jump operators
as
\begin{align}
  \sum_j \sigma_j^+ S_+^\ell S_-^m S_z^n \sigma_j^-
  = \sum_j S_+^\ell \sigma_j^+ \sigma_j^- S_-^m \p{S_z-1}^n
  = S_+^\ell \p{S + S_z} S_-^m
  \sum_{k=0}^n { n \choose k } \p{-1}^{n-k} S_z^k.
\end{align}
We now need to push the middle $S_z$ operator to the right, which we
can do using the commutation relation $\sp{S_z,S_\mu^m}=\mu mS_\mu$
derived in \eqref{eq:comm_z_mu_n}:
\begin{align}
  \sum_j \sigma_j^+ S_+^\ell S_-^m S_z^n \sigma_j^-
  = S_+^\ell S_-^m \p{S - m + S_z}
  \sum_{k=0}^n { n \choose k } \p{-1}^{n-k} S_z^k.
\end{align}
Considering now the last term in \eqref{eq:D_z+-}, we can expand
\begin{align}
  \f12 \sp{S_z, S_+^\ell S_-^m S_z^n}_+
  = \f12 \p{S_z S_+^\ell S_-^m + S_+^\ell S_-^m S_z} S_z^n,
\end{align}
where we can again use commutation relation in \eqref{eq:comm_z_mu_n}
to push $S_z$ to the right and get
\begin{align}
  S_z S_+^\ell S_-^m
  = S_+^\ell \p{\ell + S_z} S_-^m
  = \ell S_+^\ell S_-^m + S_+^\ell S_z S_-^m
  = S_+^\ell S_-^m \p{\ell - m + S_z},
\end{align}
which implies
\begin{align}
  \f12 \sp{S_z, S_z^\ell S_+^m S_-^n}_+
  = \f12 S_+^\ell S_-^m \p{\ell - m + 2 S_z} S_z^n
  = S_+^\ell S_-^m \p{\f12\sp{\ell-m} + S_z} S_z^n.
\end{align}
Putting everything together, we have that
\begin{align}
  \D\p{S_+^\ell S_-^m S_z^n}
  &= S_+^\ell S_-^m \sp{\p{S - m + S_z} \sum_{k=0}^n
    { n \choose k } \p{-1}^{n-k} S_z^k
    - \p{S + \f12\sp{\ell-m} + S_z} S_z^n} \\
  &= S_+^\ell S_-^m \sp{\p{S - m + S_z} \sum_{k=0}^{n-1}
    { n \choose k } \p{-1}^{n-k} S_z^k - \f12 \p{\ell+m} S_z^n}.
\end{align}

\end{document}

### 2018-08-16 soc_squeezing/cumulant_expansions.tex
fixed some errors

@@
 The evolution of any Heisenburg operator $\O$ with a Hamiltonian $H$
-is thus given by
+is given by
 \begin{align}
@@
   &&
-  \sp{S_z,S_\mu}_- = \mu S_\mu.
+  \sp{S_z,S_\mu}_- = \mu S_\mu
   \label{eq:comm_z_mu}
@@
   &&
-  \sp{S_+,S_-}_- = 2 S_z.
+  \sp{S_\mu,S_\nu}_- = 2\delta_{\mu,-\nu} \mu S_z
   \label{eq:comm_mu_nu}
@@
   = 2\mu S_\mu^n + S_\mu^2 S_z S_\mu^{n-2}
-  = S_\mu^n \p{\mu n + S_z},
+  = S_\mu^n \p{\mu n + S_z}
   \label{eq:comm_z_mu_n}
@@
 \begin{align}
-  \sigma_j^+ \sigma_j^- = \f12\p{\1_j+\sigma_j^z},
+  \sigma_j^+ \sigma_j^- = \f12\p{\1_j+\sigma_j^z}
   &&
@@
 \end{align}
-We now need to push the middle $S_z$ operator to the right, which we
-can do using the commutation relation $\sp{S_z,S_\mu^m}=\mu mS_\mu$
-derived in \eqref{eq:comm_z_mu_n}:
+We now need to push the $S_-^m$ to the left, which we can do using the
+commutation relation $\sp{S_z,S_\mu^m}=\mu mS_\mu$ derived in
+\eqref{eq:comm_z_mu_n}:
 \begin{align}
@@
 \end{align}
-where we can again use commutation relation in \eqref{eq:comm_z_mu_n}
-to push $S_z$ to the right and get
+where we can again use commutation relation
+$\sp{S_z,S_\mu^m}=\mu mS_\mu$ to push $S_z$ to the right and get
 \begin{align}
@@
   = S_+^\ell \p{\ell + S_z} S_-^m
-  = \ell S_+^\ell S_-^m + S_+^\ell S_z S_-^m
   = S_+^\ell S_-^m \p{\ell - m + S_z},
@@
 \begin{align}
-  \f12 \sp{S_z, S_z^\ell S_+^m S_-^n}_+
+  \f12 \sp{S_z, S_+^\ell S_-^m S_z^n}_+
   = \f12 S_+^\ell S_-^m \p{\ell - m + 2 S_z} S_z^n

### 2018-08-18 soc_squeezing/cumulant_expansions.tex
cleaned up some derivations made more progress in deriving commutator between S_+^m and S_-^n

@@
 \renewcommand{\O}{\mathcal{O}}
+
+\newcommand{\bmu}{{\bar\mu}}
 
@@
   = \gamma \p{S + S_z} \sum_{k=0}^{n-1}
-  { n \choose k } \p{-1}^{n-k} S_z^k,
+  \p{-1}^{n-k} { n \choose k } S_z^k,
 \end{align}
@@
 
-\subsection{Cumulant expansion}
+\subsection{Heisenberg systems of equations}
 
@@
 solutions. We will therefore solve these Hamiltonians approximately
-using cumulant expansions.  The basic idea is to find equations of
-motion for operators of interest, and at some point truncate
-high-order terms to arrive at a closed, reasonably-sized system of
-equations.  To benchmark this method, we will test it against exact
-one-axis twisting results.
+using systems of equations for Heisenberg operators.  The basic idea
+is to find equations of motion for operators of interest, and at some
+point truncate high-order terms to arrive at a closed,
+reasonably-sized system of equations.  To benchmark this method, we
+will test it against exact one-axis twisting results.
 
-In order to find equations of motion for $S_\mu$, $S_\mu S_z$, and
-$S_\mu S_\nu$ (i.e.~where $\mu,\nu\in\set{+1,-1}$), we first compute
-the commutator
+The equations of motion for $S_\mu$, $S_\mu S_z$, and $S_\mu S_\nu$
+(i.e.~where $\mu,\nu\in\set{+1,-1}$) under the OAT Hamiltonian
+$H_{\t{OAT}}=\chi S_z^2$ are determined by
+\begin{align}
+  \f{d}{dt} \p{S_+^\ell S_-^m S_z^n}
+  = i\chi \sp{S_z^2, S_+^\ell S_-^m S_z^n}_-
+  + \gamma \D\p{S_+^\ell S_-^m S_z^n},
+\end{align}
+where we can use identity \eqref{eq:comm_z_mu_nu_pow} to simplify
 \begin{align}
   \sp{S_z^2, S_+^\ell S_-^m S_z^n}_-
-  = \p{S_z^2 S_+^\ell S_-^m - S_+^\ell S_-^m S_z^2} S_z^n,
+  = \sp{S_z^2, S_+^\ell S_-^m}_- S_z^n
+  = S_+^\ell S_-^m \p{\sp{\ell-m}^2 + 2\sp{\ell-m} S_z} S_z^n,
 \end{align}
-where we can use the commutation relation
-$\sp{S_z,S_\mu^k}_-=\mu k S_\mu$ (see Appendix \ref{sec:identities})
-to push spin-$z$ operators (i.e.~$S_z$) to the right, as
-\begin{align}
-  S_z S_+^\ell S_-^m
-  = S_+^\ell \p{\ell + S_z} S_-^m
-  = \ell S_+^\ell S_-^m + S_+^\ell S_z S_-^m
-  = S_+^\ell S_-^m \p{\ell - m + S_z},
-\end{align}
-which implies
-\begin{align}
-  \sp{S_z^2, S_+^\ell S_-^m S_z^n}_-
-  = S_+^\ell S_-^m \p{\sp{\ell - m + S_z}^2 - S_z^2} S_z^n
-  = S_+^\ell S_-^m \p{\sp{\ell-m}^2 + 2\sp{\ell-m} S_z} S_z^n.
-\end{align}
-For the OAT Hamiltonian, we then have that
-\begin{align}
-  \f{d}{dt} \p{S_+^\ell S_-^m S_z^n}
-  = i\chi S_+^\ell S_-^m \p{\sp{\ell-m}^2 + 2\sp{\ell-m} S_z} S_z^n
-  + \gamma \D\p{S_+^\ell S_-^m S_z^n},
-\end{align}
-where $\D\p{S_+^\ell S_-^m S_z^n}$ is worked out in Appendix
-\ref{sec:decoherence}.  In order to compute squeezing of the OAT
-Hamiltonian in the presence of single-spin decay, we therefore need to
-solve the systems of equations defined by
+and the effect of decoherence through $\D\p{S_+^\ell S_-^m S_z^n}$ is
+worked out in Appendix \ref{sec:decoherence}.  In order to compute
+squeezing of the OAT Hamiltonian in the presence of single-spin decay,
+we therefore need to solve the systems of equations defined by
 \begin{align}
@@
   + \gamma S_\mu \sp{\p{S - \delta_{\mu,-1} + S_z} \sum_{k=0}^{n-1}
-    { n \choose k } \p{-1}^{n-k} S_z^k - \f12 S_z^n}
+    \p{-1}^{n-k} { n \choose k } S_z^k - \f12 S_z^n}
 \end{align}
@@
   \sp{\p{S - \delta_{\mu,-1} - \delta_{\nu,-1} + S_z} \sum_{k=0}^{n-1}
-    { n \choose k } \p{-1}^{n-k} S_z^k - S_z^n},
+    \p{-1}^{n-k} { n \choose k } S_z^k - S_z^n},
 \end{multline}
-for $\mu\ge\nu$.
+for $\mu\ge\nu$.  The initial conditions for these systems of equations can be determined from the expansions
+\begin{align}
+  S_\mu = \sum_{m=-S}^S \sqrt{\p{S-\mu m}\p{S+\mu m+1}} \op{m+\mu}{m},
+  &&
+  S_z^n = \sum_{m=-S}^S m^n \op{m},
+\end{align}
+and the initial state
+\begin{align}
+  \ket{X} = 2^{-N/2} \sum_{m=-S}^S { N \choose S+m }^{1/2} \ket{m},
+\end{align}
+from which we can determine
+\begin{align}
+  \bk{S_\mu S_z^n}_{t=0}
+  &= 2^{-N} \sum_{m=-S}^S m^n
+  \sp{{ N \choose S+m } { N \choose S+m+\mu }
+    \p{S-\mu m}\p{S+\mu m+1}}^{1/2} \\
+  &= 2^{-N} \sum_{m=-S}^S m^n { N \choose S+m } \abs{m - \mu S},
+\end{align}
+and
+\begin{multline}
+  \bk{S_\mu S_\nu S_z^n}_{t=0}
+  = 2^{-N} \sum_{m=-S}^S m^n
+  \sp{{ N \choose S+m } { N \choose S+m+\mu+\nu }}^{1/2} \\
+  \times \sp{\p{S-\nu\sp{m+\nu}}\p{S+\mu\sp{m+\nu}+1}
+    \p{S-\nu m}\p{S+\nu m+1}}^{1/2}.
+\end{multline}
+
+\section{One-axis twisting with a transverse field}
+
+We now consider again squeezing the initial state $\ket{X}$, but this
+time with the transverse field Hamiltonian
+\begin{align}
+  H_{\t{TVF}}
+  \equiv \chi S_z^2- \Omega S_x.
+\end{align}
+To determine the coherent evolution of the operator
+$S_\mu^m S_\nu^n S_z^n$, we now need to compute
+\begin{align}
+  \sp{H_{\t{TVF}}, S_\mu^m S_\nu^n S_z^n}_-
+  = \chi \sp{S_z^2, S_\mu^m S_\nu^n S_z^n}_-
+  - \Omega \sp{S_x, S_\mu^m S_\nu^n S_z^n}_-.
+\end{align}
+Here first commutator is identical to that for OAT, while for the
+second we need to compute
+\begin{align}
+  \sp{S_\rho, S_\mu^m S_\nu^n S_z^n}_-
+\end{align}
+for $\rho\in\set{+1,-1}$.
 
@@
 
-This text frequently makes use of various spin operator identities; we
-collect some particularly useful identities here for reference.
+This text makes ubiquitous use of various spin operator identities; we
+collect and derive particularly useful identities here for reference.
+The basic commutation relations between single-weight spin operators
+are
 \begin{align}
-  \f12 \sp{\sigma_j^z,\sigma_k^\mu}_- = \delta_{jk} \mu \sigma_j^\mu,
+  \f12 \sp{\sigma_j^z,\sigma_k^\mu}_-
+  &= \delta_{jk} \mu \sigma_j^\mu,
+  &
+  \sp{S_z,\sigma_j^\mu}_-
+  &= \f12 \sp{\sigma_j^z,S_\mu}_- = \mu \sigma_j^\mu,
+  &
+  \sp{S_z,S_\mu}_-
+  &= \mu S_\mu,
+  \label{eq:comm_z_mu} \\
+  \sp{\sigma_j^\mu,\sigma_k^\nu}_-
+  &= \delta_{jk} \delta_{\mu,-\nu} \mu \sigma_j^z,
+  &
+  \sp{S_\mu,\sigma_j^\nu}_-
+  &= \sp{\sigma_j^\mu,S_\nu}_-
+  = \delta_{\mu,-\nu} \mu \sigma_j^z,
+  &
+  \sp{S_\mu,S_\nu}_-
+  &= \delta_{\mu,-\nu} 2 \mu S_z.
+  \label{eq:comm_mu_nu}
+\end{align}
+These relations can be used to inductively determine commutation
+relations between higher-weight spin operators.  In particular, we can
+find that
+\begin{align}
+  S_z S_\mu^m
+  = \p{\mu S_\mu + S_\mu S_z} S_\mu^{m-1}
+  = \mu S_\mu^m + S_\mu S_z S_\mu^{m-1}
+  = 2 \mu S_\mu^m + S_\mu S_z S_\mu^{m-2}
+  = S_\mu^m \p{\mu m + S_z},
+  \label{eq:push_z_mu}
+\end{align}
+and in turn
+\begin{align}
+  S_z S_\mu^m S_\nu^n
+  = S_\mu^m \p{\mu m + S_z} S_\nu^n
+  = S_\mu^m S_\nu^n \p{\mu m + \nu n + S_z}.
+  \label{eq:push_z_mu_nu}
+\end{align}
+Repeating this process $\ell$ times tells us
+\begin{align}
+  S_z^\ell S_\mu^m S_\nu^n
+  = S_\mu^m S_\nu^n \p{\mu m + \nu n + S_z}^\ell,
+\end{align}
+which implies
+\begin{align}
+  \sp{S_z^\ell, S_\mu^m S_\nu^n}_-
+  = S_\mu^m S_\nu^n \p{\sp{\mu m + \nu n + S_z}^\ell - S_z^\ell}
+  = S_\mu^m S_\nu^n  \sum_{k=0}^{\ell-1} {\ell \choose k}
+  \p{\mu m + \nu n}^{\ell-k} S_z^k.
+  \label{eq:comm_z_mu_nu_pow}
+\end{align}
+To determine commutation relations between high-weight operators
+transverse to the $z$ axis, we compute (with $\bmu\equiv-\mu$ for
+brevity)
+\begin{align}
+  S_\mu S_\bmu^n
+  &= \p{2\mu S_z + S_\bmu S_\mu} S_\bmu^{n-1} \\
+  &= 2\mu S_z S_\bmu^{n-1} + S_\bmu S_\mu S_\bmu^{n-1} \\
+  &= 2\mu S_z S_\bmu^{n-1}
+  + S_\bmu \sp{2\mu S_z S_\bmu^{n-2}
+    + S_\bmu S_\mu S_\bmu^{n-2}} \\
+  &= 2\mu \p{S_z S_\bmu^{n-1} + S_\bmu S_z S_\bmu ^{n-2}}
+  + S_\bmu^2 S_\mu S_\bmu^{n-2} \\
+  &= 2\mu \p{S_z S_\bmu^{n-1} + S_\bmu S_z S_\bmu ^{n-2}
+    + S_\bmu^2 S_z S_\bmu ^{n-3}} + S_\bmu^3 S_\mu S_\bmu^{n-3} \\
+  &= 2\mu \sum_{k=0}^{n-1} S_\bmu^k S_z S_\bmu^{n-k-1}
+  + S_\bmu^n S_\mu,
+\end{align}
+where
+\begin{align}
+  2 \sum_{k=0}^{n-1} S_\bmu^k S_z S_\bmu^{n-k-1}
+  = 2 \sum_{k=0}^{n-1} S_\bmu^k S_\bmu^{n-k-1} \p{\bmu\sp{n-k-1} + S_z}
+  = n S_\bmu^{n-1}\p{\bmu\sp{n-1} + 2S_z},
+\end{align}
+so
+\begin{align}
+  \sp{S_\mu, S_\bmu^n}_-
+  = n S_\bmu^{n-1}\p{\bmu\sp{n-1} + 2S_z}.
+\end{align}
+Finally, we attempt to compute
+\begin{align}
+  S_\mu^m S_\bmu^n
+  &= S_\mu^{m-1} \sp{\mu n S_\bmu^{n-1}\p{\bmu\sp{n-1} + 2S_z}
+    + S_\bmu^n S_\mu} \\
+  &= \mu n \sp{S_\mu^{m-1} S_\bmu^{n-1} \p{\bmu\sp{n-1} + 2S_z}}
+  + S_\mu^{m-1} S_\bmu^n S_\mu \\
+  &= \mu n \sum_{\ell=0}^{m-1} S_\mu^{m-\ell-1} S_\bmu^{n-1}
+  \p{\bmu\sp{n-1} + 2S_z} S_\mu^\ell + S_\bmu^n S_\mu^m \\
+  &= C_{\mu mn} + S_\bmu^n S_\mu^m,
+\end{align}
+where
+\begin{align}
+  C_{\mu mn}
+  &\equiv \mu n \sum_{\ell=0}^{m-1}
+  S_\mu^{m-\ell-1} S_\bmu^{n-1} S_\mu^\ell
+  \p{\bmu\sp{n-1} + 2\mu\ell + 2S_z} \\
+  &= n \sum_{\ell=0}^{m-1}
+  S_\mu^{m-\ell-1} S_\bmu^{n-1} S_\mu^\ell
+  \p{2\ell + 1 - n + 2\mu S_z} \\
+  &= n \sum_{\ell=0}^{m-1}
+  \p{S_\bmu^{n-1} S_\mu^{m-\ell-1} + C_{\mu,m-\ell-1,n-1}} S_\mu^\ell
+  \p{2\ell + 1 - n + 2\mu S_z} \\
+  &= D_{\mu mn}
+  + n \sum_{\ell=0}^{m-1} C_{\mu,m-\ell-1,n-1}
+  S_\mu^\ell \p{2\ell + 1 - n + 2\mu S_z},
+  \label{eq:C_mn_mu}
+\end{align}
+and
+\begin{align}
+  D_{\mu mn}
+  \equiv mn S_\bmu^{n-1} S_\mu^{m-1} \p{m - n + 2\mu S_z}.
+  \label{eq:D_mn_mu}
+\end{align}
+We generally want all spin operators to be in a standard order, so we
+now wish to push all $S_z$ operators in \eqref{eq:C_mn_mu} to the
+right.  To do so, we define
+\begin{align}
+  C_{\mu mn}^{(\ell)} \equiv C_{\mu,m-\ell,n} S_\mu^\ell,
   &&
-  \sp{S_z,\sigma_j^\mu}_- = \f12 \sp{\sigma_j^z,S_\mu}_-
-  = \mu \sigma_j^\mu,
-  &&
-  \sp{S_z,S_\mu}_- = \mu S_\mu
-  \label{eq:comm_z_mu}
+  D_{\mu mn}^{(\ell)} \equiv D_{\mu,m-\ell,n} S_\mu^\ell,
+\end{align}
+where we can write
+\begin{align}
+  D_{\mu mn}^{(\ell)}
+  &\equiv D_{\mu,m-\ell,n}S_\mu^\ell \\
+  &= \p{m-\ell}n S_\bmu^{n-1} S_\mu^{m-\ell-1}
+  \p{m-\ell-n+2\mu S_z} S_\mu^\ell \\
+  &= \p{m-\ell}n S_\bmu^{n-1} S_\mu^{m-1} \p{\ell+m-n+2\mu S_z}.
+\end{align}
+We now rearrange
+\begin{align}
+  C_{\mu mn}^{(\ell)}
+  &= D_{\mu,m-\ell,n} S_\mu^\ell + n \sum_{k=0}^{m-\ell-1}
+  C_{\mu,m-\ell-k-1,n-1} S_\mu^k \p{2k+1-n+2\mu S_z} S_\mu^\ell \\
+  &= D_{\mu mn}^{(\ell)} + n \sum_{k=0}^{m-\ell-1}
+  C_{\mu,m-\ell-k-1,n-1} S_\mu^{k+\ell} \p{2\ell+2k+1-n+2\mu S_z} \\
+  &= D_{\mu mn}^{(\ell)} + n \sum_{k=0}^{m-\ell-1}
+  C_{\mu,m-1,n-1}^{(\ell+k)} \p{2\sp{\ell+k}+1-n+2\mu S_z} \\
+  &= D_{\mu mn}^{(\ell)} + n \sum_{k=\ell}^{m-1}
+  C_{\mu,m-1,n-1}^{(k)} \p{2k+1-n+2\mu S_z},
+\end{align}
+which provides a recursive formula for
+$\sp{S_\mu^m,S_\bmu^{n}}_- = C_{\mu mn} = C_{\mu mn}^{(0)}$ in which
+all operators are in a standard order.  While solving this formula may
+be more trouble than it's worth right now, to derive a system of
+equations for the two-axis twisting Hamiltonian we need to know
+$\sp{S_\mu^2,S_\bmu^n}_-$, so we expand
+\begin{align}
+  \sp{S_\mu^2,S_\bmu^n}_-
+  = C_{\mu mn}
+  = D_{\mu,2,n} + n D_{\mu,1,n-1} \p{1 - n + 2\mu S_z}
+\end{align}
+
+\vspace{3cm}
+
+\begin{align}
+  C_{\mu mn}^{(\ell)}
+  = D_{\mu mn}^{(\ell)} + n \sum_{k=\ell}^{m-1}
+  C_{\mu,m-1,n-1}^{(k)} \p{2k+1-n+2\mu S_z}
+  = D_{\mu mn}^{(\ell)} + \sum_{p=1}^{\min\p{m,n}}
+  \mathcal{E}_{\mu mn}^{(\ell,p)}
 \end{align}
 \begin{align}
-  \sp{\sigma_j^\mu,\sigma_k^\nu}_-
-  = \delta_{jk} \delta_{\mu,-\nu} \mu \sigma_j^z,
-  &&
-  \sp{S_\mu,\sigma_j^\nu}_- = \sp{\sigma_j^\mu,S_\nu}_-
-  = \delta_{\mu,-\nu} \mu \sigma_j^z,
-  &&
-  \sp{S_\mu,S_\nu}_- = 2\delta_{\mu,-\nu} \mu S_z
-  \label{eq:comm_mu_nu}
+  \mathcal{E}_{\mu mn}^{(\ell,1)}
+  = n \sum_{k_1=\ell}^{m-1} D_{\mu,m-1,n-1}^{(k_1)}
+  \p{2k_1-\sp{n-1}+2\mu S_z}
 \end{align}
 \begin{align}
-  S_z S_\mu^n
-  = \p{\mu S_\mu + S_\mu S_z} S_\mu^{n-1}
-  = \mu S_\mu^n + S_\mu S_z S_\mu^{n-1}
-  = 2\mu S_\mu^n + S_\mu^2 S_z S_\mu^{n-2}
-  = S_\mu^n \p{\mu n + S_z}
-  \label{eq:comm_z_mu_n}
+  \mathcal{E}_{\mu mn}^{(\ell,2)}
+  = n\p{n-1} \sum_{k_1=\ell}^{m-1} \sum_{k_2=k_1}^{m-1}
+  D_{\mu,m-2,n-2}^{(k_2)}
+  \p{2k_2-\sp{n-2}+2\mu S_z} \p{2k_1-\sp{n-1}+2\mu S_z}
 \end{align}
 \begin{align}
-  \sigma_j^+ \sigma_j^- = \f12\p{\1_j+\sigma_j^z}
-  &&
-  \sum_j \sigma_j^+ \sigma_j^- = S + S_z.
+  \mathcal{E}_{\mu mn}^{(k_0,p)}
+  = 2^p n \cdots \p{n-p+1}
+  \sum_{k_1=k_0}^{m-1} \cdots \sum_{k_p=k_{p-1}}^{m-1}
+  D_{\mu,m-p,n-p}^{(k_p)}
+  \sp{f\p{k_1,1}+\mu S_z} \cdots \sp{f\p{k_p,p}+\mu S_z}
 \end{align}
-Here $S\equiv N/2$.
+where
+\begin{align}
+  f\p{k,p} = k - \f12\p{n-p}.
+\end{align}
+
+\newpage
 
@@
   \equiv \sum_j\p{\sigma_j^+\O\sigma_j^-
-    - \f12\sp{\sigma_j^+\sigma_j^-,\O}_+}
-  = \sum_j \sigma_j^+\O\sigma_j^- - S \O - \f12 \sp{S_z,\O}_+.
+    - \f12\sp{\sigma_j^+\sigma_j^-,\O}_+},
+\end{align}
+where
+\begin{align}
+  \sum_j \f12\sp{\sigma_j^+\sigma_j^-,\O}_+
+  = \f12\sp{S+S_z,\O}_+
+  = S\O + \f12\sp{S_z,\O}_+
+  = \O\p{S+S_z} + \f12\sp{S_z,\O}_-,
+\end{align}
+so
+\begin{align}
+  \D\p{\O}
+  = \sum_j \sigma_j^+\O\sigma_j^- - \O\p{S + S_z} - \f12 \sp{S_z,\O}_-.
 \end{align}
@@
   = \sum_j \sigma_j^+ S_+^\ell S_-^m S_z^n \sigma_j^-
-  - S S_+^\ell S_-^m S_z^n - \f12 \sp{S_z, S_+^\ell S_-^m S_z^n}_+.
+  - S_+^\ell S_-^m S_z^n \p{S + S_z}
+  - \f12 \sp{S_z, S_+^\ell S_-^m S_z^n}_-.
   \label{eq:D_z+-}
 \end{align}
-Considering the first term, we can use a commutation relations from
-\eqref{eq:comm_z_mu} to push through and simplify the jump operators
-as
+The last term in \eqref{eq:D_z+-} is straightforward to simplify using
+\eqref{eq:push_z_mu_nu}:
+\begin{align}
+  \f12 \sp{S_z, S_+^\ell S_-^m S_z^n}_-
+  = \f12 \sp{S_z, S_+^\ell S_-^m}_- S_z^n
+  = \f12 \p{\ell-m} S_+^\ell S_-^m S_z^n,
+\end{align}
+while the first term in \eqref{eq:D_z+-} can be written in terms of
+collective spin operators if we use
+\eqref{eq:comm_z_mu}-\eqref{eq:comm_mu_nu} to push the jump operators
+together:
 \begin{align}
@@
   = S_+^\ell \p{S + S_z} S_-^m
-  \sum_{k=0}^n { n \choose k } \p{-1}^{n-k} S_z^k.
+  \sum_{k=0}^n \p{-1}^{n-k} { n \choose k } S_z^k.
 \end{align}
-We now need to push the $S_-^m$ to the left, which we can do using the
-commutation relation $\sp{S_z,S_\mu^m}=\mu mS_\mu$ derived in
-\eqref{eq:comm_z_mu_n}:
+Finally, we need to push the middle $S_z$ to the right, which we can
+do using \eqref{eq:push_z_mu}:
 \begin{align}
@@
   = S_+^\ell S_-^m \p{S - m + S_z}
-  \sum_{k=0}^n { n \choose k } \p{-1}^{n-k} S_z^k.
-\end{align}
-Considering now the last term in \eqref{eq:D_z+-}, we can expand
-\begin{align}
-  \f12 \sp{S_z, S_+^\ell S_-^m S_z^n}_+
-  = \f12 \p{S_z S_+^\ell S_-^m + S_+^\ell S_-^m S_z} S_z^n,
-\end{align}
-where we can again use commutation relation
-$\sp{S_z,S_\mu^m}=\mu mS_\mu$ to push $S_z$ to the right and get
-\begin{align}
-  S_z S_+^\ell S_-^m
-  = S_+^\ell \p{\ell + S_z} S_-^m
-  = S_+^\ell S_-^m \p{\ell - m + S_z},
-\end{align}
-which implies
-\begin{align}
-  \f12 \sp{S_z, S_+^\ell S_-^m S_z^n}_+
-  = \f12 S_+^\ell S_-^m \p{\ell - m + 2 S_z} S_z^n
-  = S_+^\ell S_-^m \p{\f12\sp{\ell-m} + S_z} S_z^n.
+  \sum_{k=0}^n \p{-1}^{n-k} { n \choose k } S_z^k.
 \end{align}
@@
   &= S_+^\ell S_-^m \sp{\p{S - m + S_z} \sum_{k=0}^n
-    { n \choose k } \p{-1}^{n-k} S_z^k
-    - \p{S + \f12\sp{\ell-m} + S_z} S_z^n} \\
+    \p{-1}^{n-k} { n \choose k } S_z^k
+    - \p{S + S_z + \f12\sp{\ell-m}} S_z^n} \\
   &= S_+^\ell S_-^m \sp{\p{S - m + S_z} \sum_{k=0}^{n-1}
-    { n \choose k } \p{-1}^{n-k} S_z^k - \f12 \p{\ell+m} S_z^n}.
+    \p{-1}^{n-k} { n \choose k } S_z^k - \f12 \p{\ell+m} S_z^n}.
 \end{align}

### 2018-08-19 soc_squeezing/cumulant_expansions.tex
more work on the same commutator

@@
 
+\newcommand{\C}{\mathcal{C}}
 \newcommand{\D}{\mathcal{D}}
@@
 
-\section{Spin operator identities}
+\section{Basic spin operator identities}
 \label{sec:identities}
@@
 collect and derive particularly useful identities here for reference.
-The basic commutation relations between single-weight spin operators
-are
+The elementary commutation relations between single-weight spin
+operators are
 \begin{align}
@@
 \end{align}
+
+\section{Commutators between high-weight transverse spin operators}
+
 To determine commutation relations between high-weight operators
-transverse to the $z$ axis, we compute (with $\bmu\equiv-\mu$ for
-brevity)
+transverse to the decoherence (i.e.~$z$) axis, we compute (with
+$\bmu\equiv-\mu$ for brevity)
 \begin{align}
@@
   &= 2\mu \sum_{k=0}^{n-1} S_\bmu^k S_z S_\bmu^{n-k-1}
-  + S_\bmu^n S_\mu,
+  + S_\bmu^n S_\mu.
+  \label{eq:comm_mu_nu_single}
 \end{align}
-where
+While \eqref{eq:comm_mu_nu_single} gives us the commutator
+$\sp{S_\mu,S_\bmu^n}_-$, we would like to enforce an ordering on
+products of collective spin operators, which allows us to only keep
+track of high-weight operators which are linearly independent.  We
+(arbitrarily) choose to impose an ordering with all $S_z$ operators on
+the right.  This choice of ordering compels us to expand
 \begin{align}
-  2 \sum_{k=0}^{n-1} S_\bmu^k S_z S_\bmu^{n-k-1}
-  = 2 \sum_{k=0}^{n-1} S_\bmu^k S_\bmu^{n-k-1} \p{\bmu\sp{n-k-1} + S_z}
-  = n S_\bmu^{n-1}\p{\bmu\sp{n-1} + 2S_z},
+  2 \mu \sum_{k=0}^{n-1} S_\bmu^k S_z S_\bmu^{n-k-1}
+  = 2 \mu \sum_{k=0}^{n-1}
+  S_\bmu^k S_\bmu^{n-k-1} \p{\bmu\sp{n-k-1} + S_z}
+  = n S_\bmu^{n-1}\p{1 - n + 2\mu S_z},
 \end{align}
@@
   \sp{S_\mu, S_\bmu^n}_-
-  = n S_\bmu^{n-1}\p{\bmu\sp{n-1} + 2S_z}.
+  = n S_\bmu^{n-1}\p{1 - n + 2\mu S_z}.
 \end{align}
-Finally, we attempt to compute
+As the next logical step, we take on the task of computing
 \begin{align}
   S_\mu^m S_\bmu^n
-  &= S_\mu^{m-1} \sp{\mu n S_\bmu^{n-1}\p{\bmu\sp{n-1} + 2S_z}
+  &= S_\mu^{m-1} \sp{n S_\bmu^{n-1}\p{1 - n + 2\mu S_z}
     + S_\bmu^n S_\mu} \\
-  &= \mu n \sp{S_\mu^{m-1} S_\bmu^{n-1} \p{\bmu\sp{n-1} + 2S_z}}
+  &= n \sp{S_\mu^{m-1} S_\bmu^{n-1} \p{1 - n + 2\mu S_z}}
   + S_\mu^{m-1} S_\bmu^n S_\mu \\
-  &= \mu n \sum_{\ell=0}^{m-1} S_\mu^{m-\ell-1} S_\bmu^{n-1}
-  \p{\bmu\sp{n-1} + 2S_z} S_\mu^\ell + S_\bmu^n S_\mu^m \\
-  &= C_{\mu mn} + S_\bmu^n S_\mu^m,
+  &= n \sum_{\ell=0}^{m-1} S_\mu^{m-\ell-1} S_\bmu^{n-1}
+  \p{1 - n + 2\mu S_z} S_\mu^\ell + S_\bmu^n S_\mu^m,
 \end{align}
-where
+which implies
+\begin{align}
+  \sp{S_\mu^m, S_\bmu^n}_-
+  = C_{\mu mn}
+  \equiv n \sum_{\ell=0}^{m-1} S_\mu^{m-\ell-1} S_\bmu^{n-1}
+  \p{1 - n + 2\mu S_z} S_\mu^\ell.
+\end{align}
+We now need rearrange the operators in $C_{\mu mn}$ into a standard
+order, which means pushing all $S_z$ operators to the right and, for
+the purposes of this calculation, all $S_\bmu$ operators on the left.
+We begin by pushing $S_\mu^\ell$ to the left of $S_z$, which takes
+$2\mu S_z\to 2\mu S_z+2\ell$, and then push $S_\mu^{m-\ell-1}$ to the
+right of $S_\bmu^{n-1}$, giving us
 \begin{align}
   C_{\mu mn}
-  &\equiv \mu n \sum_{\ell=0}^{m-1}
-  S_\mu^{m-\ell-1} S_\bmu^{n-1} S_\mu^\ell
-  \p{\bmu\sp{n-1} + 2\mu\ell + 2S_z} \\
-  &= n \sum_{\ell=0}^{m-1}
-  S_\mu^{m-\ell-1} S_\bmu^{n-1} S_\mu^\ell
-  \p{2\ell + 1 - n + 2\mu S_z} \\
   &= n \sum_{\ell=0}^{m-1}
@@
   &= D_{\mu mn}
-  + n \sum_{\ell=0}^{m-1} C_{\mu,m-\ell-1,n-1}
+  + n \sum_{\ell=0}^{m-2} C_{\mu,m-\ell-1,n-1}
   S_\mu^\ell \p{2\ell + 1 - n + 2\mu S_z},
-  \label{eq:C_mn_mu}
+  \label{eq:C_mn}
 \end{align}
-and
+where we have dropped the last term in the final sum because
+$C_{\mu,m-\ell-1,n-1}=\sp{S_\mu^{m-\ell-1},S_\bmu^{n-1}}_-=0$ if
+$\ell=m-1$, and
 \begin{align}
@@
   \equiv mn S_\bmu^{n-1} S_\mu^{m-1} \p{m - n + 2\mu S_z}.
-  \label{eq:D_mn_mu}
+  \label{eq:D_mn}
 \end{align}
-We generally want all spin operators to be in a standard order, so we
-now wish to push all $S_z$ operators in \eqref{eq:C_mn_mu} to the
-right.  To do so, we define
+To our despair, we have arrived in \eqref{eq:C_mn} to a {\it
+  recursive} expression for $C_{\mu mn}$, and we have not even managed
+to order the spin operators, as $C_{\mu,m-\ell-1,n-1}$ contains $S_z$
+operators which are to the left of $S_\mu^\ell$.  To sort all spin
+operators once and for all, we define
 \begin{align}
@@
 \end{align}
-where we can write
+which we can expand as
 \begin{align}
@@
   \p{m-\ell-n+2\mu S_z} S_\mu^\ell \\
-  &= \p{m-\ell}n S_\bmu^{n-1} S_\mu^{m-1} \p{\ell+m-n+2\mu S_z}.
+  &= \p{m-\ell}n S_\bmu^{n-1} S_\mu^{m-1} \p{\ell+m-n+2\mu S_z},
+  \label{eq:D_mn_ell}
 \end{align}
-We now rearrange
+and
 \begin{align}
   C_{\mu mn}^{(\ell)}
-  &= D_{\mu,m-\ell,n} S_\mu^\ell + n \sum_{k=0}^{m-\ell-1}
+  &= D_{\mu,m-\ell,n} S_\mu^\ell + n \sum_{k=0}^{m-\ell-2}
   C_{\mu,m-\ell-k-1,n-1} S_\mu^k \p{2k+1-n+2\mu S_z} S_\mu^\ell \\
-  &= D_{\mu mn}^{(\ell)} + n \sum_{k=0}^{m-\ell-1}
-  C_{\mu,m-\ell-k-1,n-1} S_\mu^{k+\ell} \p{2\ell+2k+1-n+2\mu S_z} \\
-  &= D_{\mu mn}^{(\ell)} + n \sum_{k=0}^{m-\ell-1}
-  C_{\mu,m-1,n-1}^{(\ell+k)} \p{2\sp{\ell+k}+1-n+2\mu S_z} \\
-  &= D_{\mu mn}^{(\ell)} + n \sum_{k=\ell}^{m-1}
-  C_{\mu,m-1,n-1}^{(k)} \p{2k+1-n+2\mu S_z},
+  &= D_{\mu mn}^{(\ell)} + n \sum_{k=0}^{m-\ell-2}
+  C_{\mu,m-\ell-k-1,n-1} S_\mu^{k+\ell} \p{2k+2\ell+1-n+2\mu S_z} \\
+  &= D_{\mu mn}^{(\ell)} + n \sum_{k=0}^{m-\ell-2}
+  C_{\mu,m-1,n-1}^{(\ell+k)} \p{2\sp{k+\ell}+1-n+2\mu S_z} \\
+  &= D_{\mu mn}^{(\ell)} + n \sum_{k=\ell}^{m-2}
+  C_{\mu,m-1,n-1}^{(k)} \p{2k+1-n+2\mu S_z}.
+  \label{eq:C_mn_ell}
 \end{align}
-which provides a recursive formula for
-$\sp{S_\mu^m,S_\bmu^{n}}_- = C_{\mu mn} = C_{\mu mn}^{(0)}$ in which
-all operators are in a standard order.  While solving this formula may
-be more trouble than it's worth right now, to derive a system of
-equations for the two-axis twisting Hamiltonian we need to know
-$\sp{S_\mu^2,S_\bmu^n}_-$, so we expand
+While the resulting expression strongly resembles that in
+\eqref{eq:C_mn}, there is one crucial difference: all spin operators
+in \eqref{eq:C_mn_ell} have been ordered.  We can now repeatedly
+substitute $C_{\mu mn}^{(\ell)}$ into itself, each time decreasing $m$
+and $n$ by 1, until one of $m$ or $n$ reaches zero.  Such repeated
+substitution yields the expansion
+\begin{align}
+  C_{\mu mn}^{(\ell)}
+  = D_{\mu mn}^{(\ell)}
+  + \sum_{p=1}^{\min\set{m,n}-1} E_{\mu mn}^{(\ell,p)},
+\end{align}
+where the first terms in the sum are
+\begin{align}
+  E_{\mu mn}^{(\ell,1)}
+  = n \sum_{k=\ell}^{m-2} D_{\mu,m-1,n-1}^{(k)}
+  \p{2k+1-n+2\mu S_z},
+\end{align}
+\begin{align}
+  E_{\mu mn}^{(\ell,2)}
+  = n \sum_{k_1=\ell}^{m-2} \p{n-1} \sum_{k_2=k_1}^{m-3}
+  D_{\mu,m-2,n-2}^{(k_2)}
+  \p{2k_2+2-n+2\mu S_z} \p{2k_1+1-n+2\mu S_z},
+\end{align}
+and more generally
+\begin{align}
+  E_{\mu mn}^{(k_0,p)}
+  = n^{\underline{p}} \sum_{\p{\v k,k_0,m,p}}
+  D_{\mu,m-p,n-p}^{(k_p)} \prod_{j=1}^p \p{2k_j+j-n+2\mu S_z}.
+  \label{eq:E_mn_kp}
+\end{align}
+with
+$n^{\underline{p}}\equiv
+n\p{n-1}\cdots\p{n-p+1}=\prod_{j=0}^{p-1}\p{n-j}$ and
+\begin{align}
+  \sum_{\p{\v k,k_0,m,p}}
+  \equiv \sum_{k_1=k_0}^{m-2} \cdots \sum_{k_p=k_{p-1}}^{m-p-1}.
+\end{align}
+Using \eqref{eq:D_mn_ell} to substitute $D_{\mu,m-p,n-p}^{(k_p)}$ into
+\eqref{eq:E_mn_kp} gives us
+\begin{align}
+  E_{\mu mn}^{(k_0,p)}
+  &= n^{\underline{p}} S_\bmu^{n-p-1} S_\mu^{m-p-1} \tag*{} \\
+  &\qquad \times \sum_{\p{\v k,k_0,m,p}}
+  \p{m-p-k_p} \p{n-p} \p{k_p+m-n+2\mu S_z} \tag*{} \\
+  &\qquad \times \prod_{j=1}^p \p{2k_j+j-n+2\mu S_z}.
+\end{align}
+We now wish to pick off the coefficients of each operator in
+$E_{\mu mn}^{\p{k_0,p}}$, so we express
+\begin{align}
+  \prod_{j=1}^p \p{2k_j+p-n+2\mu S_z}
+  = \sum_{q=0}^p \p{2\mu S_z}^q
+  \sum_{c\in\C\p{q,p}} \prod_{\alpha\notin c} \p{2k_\alpha+\alpha-n},
+  \label{eq:product_expansion}
+\end{align}
+where implicitly $\alpha\in\set{1,2,\cdots,p}$, and $\C\p{q,p}$
+denotes the set of $q$-combinations (i.e.~subsets with $q$ elements)
+of $\set{1,2,\cdots,p}$; the number of elements in $\C\p{q,p}$ is
+${p \choose q}$.  Note that for the expansion in
+\eqref{eq:product_expansion} to hold, we require that the empty
+product
+$\prod_{\alpha\notin\set{1,\cdots,p}}\p{2k_\alpha+\alpha-n}=1$.  We
+can now write
+\begin{align}
+  E_{\mu mn}^{\p{k_0,p}}
+  &= n^{\underline{p}} S_\bmu^{n-p-1} S_\mu^{m-p-1}
+  \sum_{q=0}^p \p{2\mu}^q S_z^q \tag*{} \\
+  &\qquad \times \sum_{c\in\C\p{q,p}} \sum_{\p{\v k,k_0,m,p}}
+  \p{m-p-k_p} \p{n-p} \p{k_p+m-n+2\mu S_z}
+  \prod_{\alpha\notin c} \p{2k_\alpha+\alpha-n}.
+\end{align}
+
+\vspace{3cm}
+
+Solving \eqref{eq:C_mn_ell} may be more trouble than it's worth
+right now, but to derive a system of equations for the two-axis
+twisting Hamiltonian we need to know $\sp{S_\mu^2,S_\bmu^n}_-$, so we
+expand
 \begin{align}
@@
   = D_{\mu,2,n} + n D_{\mu,1,n-1} \p{1 - n + 2\mu S_z}
-\end{align}
-
-\vspace{3cm}
-
-\begin{align}
-  C_{\mu mn}^{(\ell)}
-  = D_{\mu mn}^{(\ell)} + n \sum_{k=\ell}^{m-1}
-  C_{\mu,m-1,n-1}^{(k)} \p{2k+1-n+2\mu S_z}
-  = D_{\mu mn}^{(\ell)} + \sum_{p=1}^{\min\p{m,n}}
-  \mathcal{E}_{\mu mn}^{(\ell,p)}
-\end{align}
-\begin{align}
-  \mathcal{E}_{\mu mn}^{(\ell,1)}
-  = n \sum_{k_1=\ell}^{m-1} D_{\mu,m-1,n-1}^{(k_1)}
-  \p{2k_1-\sp{n-1}+2\mu S_z}
-\end{align}
-\begin{align}
-  \mathcal{E}_{\mu mn}^{(\ell,2)}
-  = n\p{n-1} \sum_{k_1=\ell}^{m-1} \sum_{k_2=k_1}^{m-1}
-  D_{\mu,m-2,n-2}^{(k_2)}
-  \p{2k_2-\sp{n-2}+2\mu S_z} \p{2k_1-\sp{n-1}+2\mu S_z}
-\end{align}
-\begin{align}
-  \mathcal{E}_{\mu mn}^{(k_0,p)}
-  = 2^p n \cdots \p{n-p+1}
-  \sum_{k_1=k_0}^{m-1} \cdots \sum_{k_p=k_{p-1}}^{m-1}
-  D_{\mu,m-p,n-p}^{(k_p)}
-  \sp{f\p{k_1,1}+\mu S_z} \cdots \sp{f\p{k_p,p}+\mu S_z}
-\end{align}
-where
-\begin{align}
-  f\p{k,p} = k - \f12\p{n-p}.
 \end{align}

### 2018-08-20 soc_squeezing/cumulant_expansions.tex
solved last commutator!

@@
 \renewcommand{\O}{\mathcal{O}}
+\renewcommand{\S}{\mathcal{S}}
+
+\newcommand{\Z}{\mathbb{Z}}
 
@@
   + S_\mu^{m-1} S_\bmu^n S_\mu \\
-  &= n \sum_{\ell=0}^{m-1} S_\mu^{m-\ell-1} S_\bmu^{n-1}
-  \p{1 - n + 2\mu S_z} S_\mu^\ell + S_\bmu^n S_\mu^m,
+  &= n \sum_{k=0}^{m-1} S_\mu^{m-k-1} S_\bmu^{n-1}
+  \p{1 - n + 2\mu S_z} S_\mu^k + S_\bmu^n S_\mu^m,
 \end{align}
@@
   = C_{\mu mn}
-  \equiv n \sum_{\ell=0}^{m-1} S_\mu^{m-\ell-1} S_\bmu^{n-1}
-  \p{1 - n + 2\mu S_z} S_\mu^\ell.
+  \equiv n \sum_{k=0}^{m-1} S_\mu^{m-k-1} S_\bmu^{n-1}
+  \p{1 - n + 2\mu S_z} S_\mu^k.
 \end{align}
@@
 the purposes of this calculation, all $S_\bmu$ operators on the left.
-We begin by pushing $S_\mu^\ell$ to the left of $S_z$, which takes
-$2\mu S_z\to 2\mu S_z+2\ell$, and then push $S_\mu^{m-\ell-1}$ to the
-right of $S_\bmu^{n-1}$, giving us
+We begin by pushing $S_\mu^k$ to the left of $S_z$, which takes
+$2\mu S_z\to 2\mu S_z+2k$, and then push $S_\mu^{m-k-1}$ to the right
+of $S_\bmu^{n-1}$, giving us
 \begin{align}
   C_{\mu mn}
-  &= n \sum_{\ell=0}^{m-1}
-  \p{S_\bmu^{n-1} S_\mu^{m-\ell-1} + C_{\mu,m-\ell-1,n-1}} S_\mu^\ell
-  \p{2\ell + 1 - n + 2\mu S_z} \\
+  &= n \sum_{k=0}^{m-1}
+  \p{S_\bmu^{n-1} S_\mu^{m-k-1} + C_{\mu,m-k-1,n-1}} S_\mu^k
+  \p{2k + 1 - n + 2\mu S_z} \\
   &= D_{\mu mn}
-  + n \sum_{\ell=0}^{m-2} C_{\mu,m-\ell-1,n-1}
-  S_\mu^\ell \p{2\ell + 1 - n + 2\mu S_z},
+  + n \sum_{k=0}^{m-2} C_{\mu,m-k-1,n-1}
+  S_\mu^k \p{2k + 1 - n + 2\mu S_z},
   \label{eq:C_mn}
@@
 where we have dropped the last term in the final sum because
-$C_{\mu,m-\ell-1,n-1}=\sp{S_\mu^{m-\ell-1},S_\bmu^{n-1}}_-=0$ if
-$\ell=m-1$, and
+$C_{\mu,m-k-1,n-1}=\sp{S_\mu^{m-k-1},S_\bmu^{n-1}}_-=0$ if $k=m-1$,
+and
 \begin{align}
@@
   recursive} expression for $C_{\mu mn}$, and we have not even managed
-to order the spin operators, as $C_{\mu,m-\ell-1,n-1}$ contains $S_z$
-operators which are to the left of $S_\mu^\ell$.  To sort all spin
+to order the spin operators, as $C_{\mu,m-k-1,n-1}$ contains $S_z$
+operators which are to the left of $S_\mu^k$.  To sort all spin
 operators once and for all, we define
 \begin{align}
-  C_{\mu mn}^{(\ell)} \equiv C_{\mu,m-\ell,n} S_\mu^\ell,
+  C_{\mu mn}^{(k)} \equiv C_{\mu,m-k,n} S_\mu^k,
   &&
-  D_{\mu mn}^{(\ell)} \equiv D_{\mu,m-\ell,n} S_\mu^\ell,
+  D_{\mu mn}^{(k)} \equiv D_{\mu,m-k,n} S_\mu^k,
 \end{align}
@@
 \begin{align}
-  D_{\mu mn}^{(\ell)}
-  &\equiv D_{\mu,m-\ell,n}S_\mu^\ell \\
-  &= \p{m-\ell}n S_\bmu^{n-1} S_\mu^{m-\ell-1}
-  \p{m-\ell-n+2\mu S_z} S_\mu^\ell \\
-  &= \p{m-\ell}n S_\bmu^{n-1} S_\mu^{m-1} \p{\ell+m-n+2\mu S_z},
-  \label{eq:D_mn_ell}
+  D_{\mu mn}^{(k)}
+  &\equiv D_{\mu,m-k,n}S_\mu^k \\
+  &= \p{m-k}n S_\bmu^{n-1} S_\mu^{m-k-1}
+  \p{m-k-n+2\mu S_z} S_\mu^k \\
+  &= \p{m-k}n S_\bmu^{n-1} S_\mu^{m-1} \p{k+m-n+2\mu S_z},
+  \label{eq:D_mn_k}
 \end{align}
@@
 \begin{align}
-  C_{\mu mn}^{(\ell)}
-  &= D_{\mu,m-\ell,n} S_\mu^\ell + n \sum_{k=0}^{m-\ell-2}
-  C_{\mu,m-\ell-k-1,n-1} S_\mu^k \p{2k+1-n+2\mu S_z} S_\mu^\ell \\
-  &= D_{\mu mn}^{(\ell)} + n \sum_{k=0}^{m-\ell-2}
-  C_{\mu,m-\ell-k-1,n-1} S_\mu^{k+\ell} \p{2k+2\ell+1-n+2\mu S_z} \\
-  &= D_{\mu mn}^{(\ell)} + n \sum_{k=0}^{m-\ell-2}
-  C_{\mu,m-1,n-1}^{(\ell+k)} \p{2\sp{k+\ell}+1-n+2\mu S_z} \\
-  &= D_{\mu mn}^{(\ell)} + n \sum_{k=\ell}^{m-2}
-  C_{\mu,m-1,n-1}^{(k)} \p{2k+1-n+2\mu S_z}.
-  \label{eq:C_mn_ell}
+  C_{\mu mn}^{(k)}
+  &= D_{\mu,m-k,n} S_\mu^k + n \sum_{j=0}^{m-k-2}
+  C_{\mu,m-k-j-1,n-1} S_\mu^j \p{2j+1-n+2\mu S_z} S_\mu^k \\
+  &= D_{\mu mn}^{(k)} + n \sum_{j=0}^{m-k-2}
+  C_{\mu,m-k-j-1,n-1} S_\mu^{j+k} \p{2j+2k+1-n+2\mu S_z} \\
+  &= D_{\mu mn}^{(k)} + n \sum_{j=0}^{m-k-2}
+  C_{\mu,m-1,n-1}^{(k+j)} \p{2\sp{j+k}+1-n+2\mu S_z} \\
+  &= D_{\mu mn}^{(k)} + n \sum_{j=k}^{m-2}
+  C_{\mu,m-1,n-1}^{(j)} \p{2j+1-n+2\mu S_z}.
+  \label{eq:C_mn_k}
 \end{align}
@@
 \eqref{eq:C_mn}, there is one crucial difference: all spin operators
-in \eqref{eq:C_mn_ell} have been ordered.  We can now repeatedly
-substitute $C_{\mu mn}^{(\ell)}$ into itself, each time decreasing $m$
+in \eqref{eq:C_mn_k} have been ordered.  We can now repeatedly
+substitute $C_{\mu mn}^{(k)}$ into itself, each time decreasing $m$
 and $n$ by 1, until one of $m$ or $n$ reaches zero.  Such repeated
@@
 \begin{align}
-  C_{\mu mn}^{(\ell)}
-  = D_{\mu mn}^{(\ell)}
-  + \sum_{p=1}^{\min\set{m,n}-1} E_{\mu mn}^{(\ell,p)},
+  C_{\mu mn}
+  = C_{\mu mn}^{(0)}
+  = D_{\mu mn}
+  + \sum_{p=1}^{\min\set{m,n}-1} E_{\mu mn}^{(p)},
 \end{align}
-where the first terms in the sum are
+where the first terms in the sum over $p$ is
 \begin{align}
-  E_{\mu mn}^{(\ell,1)}
-  = n \sum_{k=\ell}^{m-2} D_{\mu,m-1,n-1}^{(k)}
+  E_{\mu mn}^{(1)}
+  = n \sum_{k=0}^{m-2} D_{\mu,m-1,n-1}^{(k)}
   \p{2k+1-n+2\mu S_z},
+  \label{eq:E_mn_1_first}
 \end{align}
+the second
 \begin{align}
-  E_{\mu mn}^{(\ell,2)}
-  = n \sum_{k_1=\ell}^{m-2} \p{n-1} \sum_{k_2=k_1}^{m-3}
+  E_{\mu mn}^{(2)}
+  = n \sum_{k_1=0}^{m-2} \p{n-1} \sum_{k_2=k_1}^{m-3}
   D_{\mu,m-2,n-2}^{(k_2)}
@@
 \end{align}
-and more generally
+and more generally for $p>1$,
 \begin{align}
-  E_{\mu mn}^{(k_0,p)}
-  = n^{\underline{p}} \sum_{\p{\v k,k_0,m,p}}
-  D_{\mu,m-p,n-p}^{(k_p)} \prod_{j=1}^p \p{2k_j+j-n+2\mu S_z}.
-  \label{eq:E_mn_kp}
+  E_{\mu mn}^{(p)}
+  = n^{\underline{p}}
+  \sum_{k_1=0}^{m-2}\cdots\sum_{k_p=k_{p-1}}^{m-p-1}
+  D_{\mu,m-p,n-p}^{(k_p)} \prod_{j=1}^p \p{2k_j+j-n+2\mu S_z},
 \end{align}
-with
-$n^{\underline{p}}\equiv
-n\p{n-1}\cdots\p{n-p+1}=\prod_{j=0}^{p-1}\p{n-j}$ and
+with $n^{\underline{p}}\equiv n\p{n-1}\cdots\p{n-p+1}$.  Observing
+that we always have $k_1\le k_2\le\cdots\le k_p$, we can rearrange the
+sums to get that for $p>1$
 \begin{align}
-  \sum_{\p{\v k,k_0,m,p}}
-  \equiv \sum_{k_1=k_0}^{m-2} \cdots \sum_{k_p=k_{p-1}}^{m-p-1}.
+  E_{\mu mn}^{(p)}
+  = n^{\underline{p}}
+  \sum_{\ell=0}^{m-p-1} D_{\mu,m-p,n-p}^{(\ell)} \p{2\ell+p-n+2\mu S_z}
+  \sum_{\substack{\v k\in\Z_{\ell+1}^{p-1}\\k_1\le k_2\le\cdots\le k_{p-1}}}
+  \prod_{j=1}^{p-1} \p{2k_j+j-n+2\mu S_z},
+  \label{eq:E_mn_p_semifinal}
 \end{align}
-Using \eqref{eq:D_mn_ell} to substitute $D_{\mu,m-p,n-p}^{(k_p)}$ into
-\eqref{eq:E_mn_kp} gives us
+where $\Z_q$ denotes the set of $q$ smallest non-negative integers,
+e.g.~$\Z_{\ell+1}=\set{0,1,\cdots,\ell}$.  We now expand
 \begin{align}
-  E_{\mu mn}^{(k_0,p)}
-  &= n^{\underline{p}} S_\bmu^{n-p-1} S_\mu^{m-p-1} \tag*{} \\
-  &\qquad \times \sum_{\p{\v k,k_0,m,p}}
-  \p{m-p-k_p} \p{n-p} \p{k_p+m-n+2\mu S_z} \tag*{} \\
-  &\qquad \times \prod_{j=1}^p \p{2k_j+j-n+2\mu S_z}.
-\end{align}
-We now wish to pick off the coefficients of each operator in
-$E_{\mu mn}^{\p{k_0,p}}$, so we express
-\begin{align}
-  \prod_{j=1}^p \p{2k_j+p-n+2\mu S_z}
-  = \sum_{q=0}^p \p{2\mu S_z}^q
-  \sum_{c\in\C\p{q,p}} \prod_{\alpha\notin c} \p{2k_\alpha+\alpha-n},
+  \prod_{j=1}^{p-1} \p{2k_j+j-n+2\mu S_z}
+  = \sum_{q=0}^{p-1} \p{2\mu S_z}^q
+  \sum_{c\in\C_{p-1-q}\p{\Z_{p-1}}} \prod_{j\in c} \p{2k_{j+1}+j+1-n},
   \label{eq:product_expansion}
 \end{align}
-where implicitly $\alpha\in\set{1,2,\cdots,p}$, and $\C\p{q,p}$
-denotes the set of $q$-combinations (i.e.~subsets with $q$ elements)
-of $\set{1,2,\cdots,p}$; the number of elements in $\C\p{q,p}$ is
-${p \choose q}$.  Note that for the expansion in
-\eqref{eq:product_expansion} to hold, we require that the empty
-product
-$\prod_{\alpha\notin\set{1,\cdots,p}}\p{2k_\alpha+\alpha-n}=1$.  We
-can now write
+where $\C\p{k,\S}$ denotes the set of $k$-combinations (i.e.~subsets
+with $k$ elements) of $\S$; the number of elements in
+$\C_{p-1-q}\p{\Z_{p-1}}$ is
+${ p-1 \choose p-1-q } = { p-1 \choose q }$.  Substituting
+\eqref{eq:D_mn_k} and \eqref{eq:product_expansion} into
+\eqref{eq:E_mn_p_semifinal}, we get that for $p>1$,
 \begin{align}
-  E_{\mu mn}^{\p{k_0,p}}
-  &= n^{\underline{p}} S_\bmu^{n-p-1} S_\mu^{m-p-1}
-  \sum_{q=0}^p \p{2\mu}^q S_z^q \tag*{} \\
-  &\qquad \times \sum_{c\in\C\p{q,p}} \sum_{\p{\v k,k_0,m,p}}
-  \p{m-p-k_p} \p{n-p} \p{k_p+m-n+2\mu S_z}
-  \prod_{\alpha\notin c} \p{2k_\alpha+\alpha-n}.
+  E_{\mu mn}^{(p)}
+  = n^{\underline{p}} S_\bmu^{n-p-1} S_\mu^{m-p-1}
+  \sum_{s=0}^2 f_{mnp\ell}^{(s)} \p{2\mu S_z}^s
+  \sum_{q=0}^{p-1} g_{np\ell q} \p{2\mu S_z}^q,
+  \label{eq:E_mn_p}
+\end{align}
+where
+\begin{align}
+  f_{mnp\ell}^{(0)}
+  &= \p{m-p-\ell} \p{n-p} \p{\ell+m-n} \p{2\ell-n+p}, \\
+  f_{mnp\ell}^{(1)}
+  &= \p{m-p-\ell} \p{n-p} \p{3\ell+m-2n+p}, \\
+  f_{mnp\ell}^{(2)}
+  &= \p{m-p-\ell} \p{n-p}, \\
+  g_{np\ell q}
+  &=
+  \sum_{c\in\C_{p-1-q}\p{\Z_{p-1}}}
+  \sum_{\substack{\v k\in\Z_{\ell+1}^{p-1}\\k_1\le k_2\le\cdots\le k_{p-1}}}
+  \prod_{j\in c} \p{2k_{j+1}+j+1-n},
+\end{align}
+and for $p=1$ we can resolve \eqref{eq:E_mn_1_first} to get
+\begin{align}
+  E_{\mu mn}^{(1)}
+  = \f16 mn \p{m-1} \p{3n-m-1} S_\bmu^{n-1} S_\mu ^{m-1} \p{m+2\mu S_z}.
+  \label{eq:E_mn_1}
+\end{align}
+With reference to \eqref{eq:E_mn_p}-\eqref{eq:E_mn_1}, we finally have
+\begin{align}
+  \sp{S_\mu^m, S_\bmu^n}_-
+  = C_{\mu mn}
+  = mn S_\bmu^{n-1} S_\mu^{m-1}
+  \sum_{q\in\set{0,1}} \p{m-n}^{1-q}\p{2\mu S_z}^q
+  + \sum_{p=1}^{\min\set{m,n}-1} E_{\mu mn}^{(p)}.
 \end{align}
@@
 
-Solving \eqref{eq:C_mn_ell} may be more trouble than it's worth
-right now, but to derive a system of equations for the two-axis
-twisting Hamiltonian we need to know $\sp{S_\mu^2,S_\bmu^n}_-$, so we
-expand
+To derive a system of equations for the two-axis twisting Hamiltonian
+we need to know $\sp{S_\mu^2,S_\bmu^n}_-$, so we expand
 \begin{align}

### 2018-08-20 soc_squeezing/cumulant_expansions.tex
changed strategy for simplifying last commutator in Appendix A

@@
 that we always have $k_1\le k_2\le\cdots\le k_p$, we can rearrange the
-sums to get that for $p>1$
+sums to get that
 \begin{align}
@@
   \sum_{\ell=0}^{m-p-1} D_{\mu,m-p,n-p}^{(\ell)} \p{2\ell+p-n+2\mu S_z}
-  \sum_{\substack{\v k\in\Z_{\ell+1}^{p-1}\\k_1\le k_2\le\cdots\le k_{p-1}}}
+  \sum_{\substack{\v k\in\Z_{\ell+1}^{p-1}\\k_1\le\cdots\le k_{p-1}}}
   \prod_{j=1}^{p-1} \p{2k_j+j-n+2\mu S_z},
-  \label{eq:E_mn_p_semifinal}
 \end{align}
 where $\Z_q$ denotes the set of $q$ smallest non-negative integers,
-e.g.~$\Z_{\ell+1}=\set{0,1,\cdots,\ell}$.  We now expand
+e.g.~$\Z_{\ell+1}=\set{0,1,\cdots,\ell}$.  Defining for shorthand
+\begin{align}
+  \sum_{\p{\v k,p,\ell}} X \equiv
+  \sum_{\substack{\v k\in\Z_{\ell+1}^{p-1}\\k_1\le\cdots\le k_{p-1}}} X,
+  &&
+  F_{\mu np} \equiv p - n + 2\mu S_z,
+\end{align}
+we can say
+\begin{align}
+  E_{\mu mn}^{(p)}
+  = n^{\underline{p}}
+  \sum_{\ell=0}^{m-p-1} D_{\mu,m-p,n-p}^{(\ell)} \p{2\ell+F_{\mu np}}
+  \sum_{\p{\v k,p,\ell}} \prod_{j=1}^{p-1} \p{2k_{p-j}-j+F_{\mu np}}.
+  \label{eq:E_mn_p_k_sum}
+\end{align}
+We now define
+\begin{align}
+  f_{\mu np\ell}\p{k,q} \equiv \p{\ell-k+q} \p{\ell+k-q+F_{\mu np}},
+\end{align}
+and simplify
+\begin{align}
+  \sum_{\p{\v k,p,\ell}} \prod_{j=1}^{p-1} \p{2k_{p-j}-j+F_{\mu np}}
+  &= \sum_{\p{\v k;p-1;\ell}}
+  \prod_{j=2}^{p-1} \p{2k_{p-j}-j+F_{\mu np}}
+  f_{\mu n\ell}\p{k_{p-2},1} \\
+  &= \f12 \sum_{\p{\v k,p-2,\ell}}
+  \prod_{j=3}^{p-1} \p{2k_{p-j}-j+F_{\mu np}}
+  f_{\mu np\ell}\p{k_{p-3},1} f_{\mu np\ell}\p{k_{p-3},2} \\
+  &\qquad \vdots \tag*{} \\
+  &= \f1{r!} \sum_{\p{\v k,p-r,\ell}}
+  \prod_{j=1+r}^{p-1} \p{2k_{p-j}-j+F_{\mu np}}
+  \prod_{q=1}^r f_{\mu np\ell}\p{k_{p-1-r},q} \\
+  &\qquad \vdots \tag*{} \\
+  &= \f1{\p{p-1}!} \prod_{q=1}^{p-1} f_{\mu np\ell}\p{0,q} \\
+  &= { \ell + p - 1 \choose p - 1 }
+  \prod_{q=1}^{p-1} \p{\ell-q+F_{\mu np}}.
+\end{align}
+Substitution into \eqref{eq:E_mn_p_k_sum} yields
+\begin{align}
+  E_{\mu mn}^{(p)}
+  = n^{\underline{p}} \sum_{\ell=0}^{m-p-1} D_{\mu,m-p,n-p}^{(\ell)}
+  { \ell + p - 1 \choose p - 1 }
+  \prod_{q=0}^{p-1} \p{\ell-q+F_{\mu np}}
+  = n^{\underline{p}} S_\bmu^{n-1} S_\mu^{m-1} G_{\mu mnp}
+\end{align}
+for
+\begin{align}
+  G_{\mu mnp}
+  \equiv \sum_{\ell=0}^{m-p-1} \p{m-p-\ell} \p{n-p}
+  \p{\ell+m-p+F_{\mu np}}
+  { \ell + p - 1 \choose p - 1 }
+  \prod_{q=0}^{p-1} \p{\ell-q+F_{\mu np}}.
+\end{align}
+
+\vspace{3cm}
+
 \begin{align}

### 2018-08-20 soc_squeezing/cumulant_expansions.tex
update to last comutator derivation

--- 2018-08-20 soc_squeezing/cumulant_expansions.tex
@@
 \end{align}
-Here first commutator is identical to that for OAT, while for the
-second we need to compute
-\begin{align}
-  \sp{S_\rho, S_\mu^m S_\nu^n S_z^n}_-
-\end{align}
-for $\rho\in\set{+1,-1}$.
+Here first commutator is identical to that for OAT, while
 
@@
   + \sum_{p=1}^{\min\set{m,n}-1} E_{\mu mn}^{(p)},
+  \label{eq:C_mn_E}
 \end{align}
@@
   \p{2k+1-n+2\mu S_z},
-  \label{eq:E_mn_1_first}
 \end{align}
@@
   E_{\mu mn}^{(p)}
-  = n^{\underline{p}}
+  = \f{n!}{\p{n-p}!}
   \sum_{k_1=0}^{m-2}\cdots\sum_{k_p=k_{p-1}}^{m-p-1}
-  D_{\mu,m-p,n-p}^{(k_p)} \prod_{j=1}^p \p{2k_j+j-n+2\mu S_z},
+  D_{\mu,m-p,n-p}^{(k_p)} \prod_{j=1}^p \p{2k_j+j-n+2\mu S_z}.
+  \label{eq:E_mn_p}
 \end{align}
-with $n^{\underline{p}}\equiv n\p{n-1}\cdots\p{n-p+1}$.  Observing
-that we always have $k_1\le k_2\le\cdots\le k_p$, we can rearrange the
-sums to get that for $p>1$
+In principle, the expressions in \eqref{eq:D_mn} and
+\eqref{eq:C_mn_E}-\eqref{eq:E_mn_p} provide a well-defined expansion
+for the commutator $\sp{S_\mu^m,S_\nu^n}_- = C_{\mu mn}$, but the
+expression for $E_{\mu mn}^{(p)}$ in \eqref{eq:E_mn_p} is -- put it
+lightly -- quite a mess, involving a sum over $p$ mutually related
+intermediate variables!  We therefore devote the rest of this section
+to simplifying $E_{\mu mn}^{(p)}$.
+
+Observing that we always have $k_1\le k_2\le\cdots\le k_p$, we can
+reverse the order of the sums in \eqref{eq:E_mn_p} and re-index
+$k_p\to\ell$ to get
 \begin{align}
   E_{\mu mn}^{(p)}
-  = n^{\underline{p}}
+  = \f{n!}{\p{n-p}!}
   \sum_{\ell=0}^{m-p-1} D_{\mu,m-p,n-p}^{(\ell)} \p{2\ell+p-n+2\mu S_z}
-  \sum_{\substack{\v k\in\Z_{\ell+1}^{p-1}\\k_1\le k_2\le\cdots\le k_{p-1}}}
+  \sum_{\substack{\v k\in\Z_{\ell+1}^{p-1}\\k_{p-1}\ge\cdots\ge k_1}}
   \prod_{j=1}^{p-1} \p{2k_j+j-n+2\mu S_z},
-  \label{eq:E_mn_p_semifinal}
 \end{align}
 where $\Z_q$ denotes the set of $q$ smallest non-negative integers,
-e.g.~$\Z_{\ell+1}=\set{0,1,\cdots,\ell}$.  We now expand
+e.g.~$\Z_{\ell+1}=\set{0,1,\cdots,\ell}$.  Defining for shorthand
 \begin{align}
-  \prod_{j=1}^{p-1} \p{2k_j+j-n+2\mu S_z}
-  = \sum_{q=0}^{p-1} \p{2\mu S_z}^q
-  \sum_{c\in\C_{p-1-q}\p{\Z_{p-1}}} \prod_{j\in c} \p{2k_{j+1}+j+1-n},
-  \label{eq:product_expansion}
+  \sum_{\p{\v k,p,\ell}} X \equiv
+  \sum_{\substack{\v k\in\Z_{\ell+1}^{p-1}\\k_{p-1}\ge\cdots\ge k_1}} X,
+  &&
+  F_{\mu np} \equiv p - n + 2\mu S_z,
 \end{align}
-where $\C\p{k,\S}$ denotes the set of $k$-combinations (i.e.~subsets
-with $k$ elements) of $\S$; the number of elements in
-$\C_{p-1-q}\p{\Z_{p-1}}$ is
-${ p-1 \choose p-1-q } = { p-1 \choose q }$.  Substituting
-\eqref{eq:D_mn_k} and \eqref{eq:product_expansion} into
-\eqref{eq:E_mn_p_semifinal}, we get that for $p>1$,
+we can say
 \begin{align}
   E_{\mu mn}^{(p)}
-  = n^{\underline{p}} S_\bmu^{n-p-1} S_\mu^{m-p-1}
-  \sum_{s=0}^2 f_{mnp\ell}^{(s)} \p{2\mu S_z}^s
-  \sum_{q=0}^{p-1} g_{np\ell q} \p{2\mu S_z}^q,
-  \label{eq:E_mn_p}
+  = \f{n!}{\p{n-p}!}
+  \sum_{\ell=0}^{m-p-1} D_{\mu,m-p,n-p}^{(\ell)} \p{2\ell+F_{\mu np}}
+  \sum_{\p{\v k,p,\ell}} \prod_{j=1}^{p-1} \p{2k_{p-j}-j+F_{\mu np}}.
+  \label{eq:E_mn_p_k_sum}
 \end{align}
-where
+We now define
 \begin{align}
-  f_{mnp\ell}^{(0)}
-  &= \p{m-p-\ell} \p{n-p} \p{\ell+m-n} \p{2\ell-n+p}, \\
-  f_{mnp\ell}^{(1)}
-  &= \p{m-p-\ell} \p{n-p} \p{3\ell+m-2n+p}, \\
-  f_{mnp\ell}^{(2)}
-  &= \p{m-p-\ell} \p{n-p}, \\
-  g_{np\ell q}
-  &=
-  \sum_{c\in\C_{p-1-q}\p{\Z_{p-1}}}
-  \sum_{\substack{\v k\in\Z_{\ell+1}^{p-1}\\k_1\le k_2\le\cdots\le k_{p-1}}}
-  \prod_{j\in c} \p{2k_{j+1}+j+1-n},
+  f_{\mu np\ell}\p{k,q} \equiv \p{\ell-k+q} \p{\ell+k-q+F_{\mu np}},
 \end{align}
-and for $p=1$ we can resolve \eqref{eq:E_mn_1_first} to get
+and evaluate sums successively over $k_{p-1},k_{p-2},\cdots,k_2,k_1$
+respectively from $k_{p-2},k_{p-3},\cdots,k_1,0$ to $\ell$, finding
 \begin{align}
-  E_{\mu mn}^{(1)}
-  = \f16 mn \p{m-1} \p{3n-m-1} S_\bmu^{n-1} S_\mu ^{m-1} \p{m+2\mu S_z}.
-  \label{eq:E_mn_1}
+  \sum_{\p{\v k,p,\ell}} \prod_{j=1}^{p-1} \p{2k_{p-j}-j+F_{\mu np}}
+  &= \sum_{\p{\v k;p-1;\ell}}
+  \prod_{j=2}^{p-1} \p{2k_{p-j}-j+F_{\mu np}}
+  f_{\mu n\ell}\p{k_{p-2},1} \\
+  &= \f12 \sum_{\p{\v k,p-2,\ell}}
+  \prod_{j=3}^{p-1} \p{2k_{p-j}-j+F_{\mu np}}
+  f_{\mu np\ell}\p{k_{p-3},1} f_{\mu np\ell}\p{k_{p-3},2} \\
+  &= \f1{r!} \sum_{\p{\v k,p-r,\ell}}
+  \prod_{j=1+r}^{p-1} \p{2k_{p-j}-j+F_{\mu np}}
+  \prod_{q=1}^r f_{\mu np\ell}\p{k_{p-1-r},q} \\
+  &= \f1{\p{p-1}!} \prod_{q=1}^{p-1} f_{\mu np\ell}\p{0,q} \\
+  &= { \ell + p - 1 \choose p - 1 }
+  \prod_{q=1}^{p-1} \p{\ell-q+F_{\mu np}}.
 \end{align}
-With reference to \eqref{eq:E_mn_p}-\eqref{eq:E_mn_1}, we finally have
+Substitution of this result together with $D_{\mu,m-p,n-p}^{(\ell)}$
+as given in using \eqref{eq:D_mn_k} into $E_{\mu mn}^{(p)}$ as
+expressed in \eqref{eq:E_mn_p_k_sum} then gives us
 \begin{align}
-  \sp{S_\mu^m, S_\bmu^n}_-
-  = C_{\mu mn}
-  = mn S_\bmu^{n-1} S_\mu^{m-1}
-  \sum_{q\in\set{0,1}} \p{m-n}^{1-q}\p{2\mu S_z}^q
-  + \sum_{p=1}^{\min\set{m,n}-1} E_{\mu mn}^{(p)}.
+  E_{\mu mn}^{(p)}
+  = \f{n!}{\p{n-p-1}!} S_\bmu^{n-p-1} S_\mu^{m-p-1} G_{\mu mnp}
+  \label{eq:E_mn_p_G}
 \end{align}
+with
+\begin{align}
+  G_{\mu mnp}
+  \equiv
+  \sum_{\ell=0}^{m-p-1} { \ell + p - 1 \choose p - 1 }
+  \p{m-p-\ell} \p{\ell+m-p+F_{\mu np}}
+  \p{2\ell + F_{\mu np}}
+  \prod_{q=1}^{p-1} \p{\ell-q+F_{\mu np}}.
+\end{align}
+
+% \vspace{3cm}
+
+% With reference to \eqref{eq:E_mn_p_G}, we finally have
+% \begin{align}
+%   \sp{S_\mu^m, S_\bmu^n}_-
+%   = C_{\mu mn}
+%   = mn S_\bmu^{n-1} S_\mu^{m-1}
+%   \sum_{q\in\set{0,1}} \p{m-n}^{1-q}\p{2\mu S_z}^q
+%   + \sum_{p=1}^{\min\set{m,n}-1} E_{\mu mn}^{(p)}.
+% \end{align}
+
+% \begin{align}
+%   \prod_{j=1}^{p-1} \p{2k_j+j-n+2\mu S_z}
+%   = \sum_{q=0}^{p-1} \p{2\mu S_z}^q
+%   \sum_{c\in\C_{p-1-q}\p{\Z_{p-1}}} \prod_{j\in c} \p{2k_{j+1}+j+1-n},
+%   \label{eq:product_expansion}
+% \end{align}
+% where $\C\p{k,\S}$ denotes the set of $k$-combinations (i.e.~subsets
+% with $k$ elements) of $\S$; the number of elements in
+% $\C_{p-1-q}\p{\Z_{p-1}}$ is
+% ${ p-1 \choose p-1-q } = { p-1 \choose q }$.
 

### 2018-08-21 soc_squeezing/cumulant_expansions.tex
finished commutator derivation

@@
 \end{align}
+Similarly finding commutation relations between high-weight transverse
+spin operators, i.e.~$S_\mu^m$ and $S_\nu^n$, turns out to be
+considerably more difficult than the cases we have worked out thus
+far.  We therefore save this case for Appendix
+\ref{sec:comm_transverse}.
 
+\newpage
 \section{Commutators between high-weight transverse spin operators}
+\label{sec:comm_transverse}
 
@@
 products of collective spin operators, which allows us to only keep
-track of high-weight operators which are linearly independent.  We
+track of high-weight operators that are linearly independent.  We
 (arbitrarily) choose to impose an ordering with all $S_z$ operators on
@@
 To our despair, we have arrived in \eqref{eq:C_mn} to a {\it
-  recursive} expression for $C_{\mu mn}$, and we have not even managed
-to order the spin operators, as $C_{\mu,m-k-1,n-1}$ contains $S_z$
-operators which are to the left of $S_\mu^k$.  To sort all spin
-operators once and for all, we define
+  recursive} expression for $C_{\mu mn}$.  Furthermore, we have not
+even managed to order the spin operators, as $C_{\mu,m-k-1,n-1}$
+contains $S_z$ operators which are to the left of $S_\mu^k$.  To sort
+all spin operators once and for all, we define
 \begin{align}
@@
 \end{align}
-While the resulting expression strongly resembles that in
-\eqref{eq:C_mn}, there is one crucial difference: all spin operators
-in \eqref{eq:C_mn_k} have been ordered.  We can now repeatedly
-substitute $C_{\mu mn}^{(k)}$ into itself, each time decreasing $m$
-and $n$ by 1, until one of $m$ or $n$ reaches zero.  Such repeated
-substitution yields the expansion
+While the resulting expression in \eqref{eq:C_mn_k} strongly resembles
+that in \eqref{eq:C_mn}, there is one crucial difference: all spin
+operators in \eqref{eq:C_mn_k} have been sorted into a standard order.
+We can now repeatedly substitute $C_{\mu mn}^{(k)}$ into itself, each
+time decreasing $m$ and $n$ by 1, until one of $m$ or $n$ reaches
+zero.  Such repeated substitution yields the expansion
 \begin{align}
@@
 In principle, the expressions in \eqref{eq:D_mn} and
-\eqref{eq:C_mn_E}-\eqref{eq:E_mn_p} provide a well-defined expansion
-for the commutator $\sp{S_\mu^m,S_\nu^n}_- = C_{\mu mn}$, but the
-expression for $E_{\mu mn}^{(p)}$ in \eqref{eq:E_mn_p} is -- put it
-lightly -- quite a mess, involving a sum over $p$ mutually related
-intermediate variables!  We therefore devote the rest of this section
-to simplifying $E_{\mu mn}^{(p)}$.
+\eqref{eq:C_mn_E}-\eqref{eq:E_mn_p} provide a complete expansion for
+the commutator $\sp{S_\mu^m,S_\nu^n}_- = C_{\mu mn}$, but the result
+is -- to put it lightly -- quite a mess: the expression for
+$E_{\mu mp}^{(p)}$ in \eqref{eq:E_mn_p} involves a sum over $p$
+mutually dependent intermediate variables!  We therefore devote the
+rest of this section to simplifying our result for the commutator
+$\sp{S_\mu^m,S_\nu^n}_-$.
 
 Observing that we always have $k_1\le k_2\le\cdots\le k_p$, we can
-reverse the order of the sums in \eqref{eq:E_mn_p} and re-index
+reverse the order of the sums in \eqref{eq:E_mn_p} and relabel
 $k_p\to\ell$ to get
@@
   \sum_{\ell=0}^{m-p-1} D_{\mu,m-p,n-p}^{(\ell)} \p{2\ell+p-n+2\mu S_z}
-  \sum_{\substack{\v k\in\Z_{\ell+1}^{p-1}\\k_{p-1}\ge\cdots\ge k_1}}
+  \sum_{\substack{\v k\in\Z_{\ell+1}^{p-1}\\
+      k_{p-2}\ge k_{p-1}\ge\cdots\ge k_1}}
   \prod_{j=1}^{p-1} \p{2k_j+j-n+2\mu S_z},
@@
   \sum_{\p{\v k,p,\ell}} X \equiv
-  \sum_{\substack{\v k\in\Z_{\ell+1}^{p-1}\\k_{p-1}\ge\cdots\ge k_1}} X,
+  \sum_{\substack{\v k\in\Z_{\ell+1}^{p-1}\\
+      k_{p-1}\ge k_{p-2}\ge\cdots\ge k_1}} X,
   &&
@@
   \sum_{\p{\v k,p,\ell}} \prod_{j=1}^{p-1} \p{2k_{p-j}-j+F_{\mu np}}.
-  \label{eq:E_mn_p_k_sum}
+  \label{eq:E_mn_p_sum}
 \end{align}
@@
 as given in using \eqref{eq:D_mn_k} into $E_{\mu mn}^{(p)}$ as
-expressed in \eqref{eq:E_mn_p_k_sum} then gives us
+expressed in \eqref{eq:E_mn_p_sum} then gives us
 \begin{align}
@@
   = \f{n!}{\p{n-p-1}!} S_\bmu^{n-p-1} S_\mu^{m-p-1} G_{\mu mnp}
-  \label{eq:E_mn_p_G}
 \end{align}
@@
   G_{\mu mnp}
-  \equiv
-  \sum_{\ell=0}^{m-p-1} { \ell + p - 1 \choose p - 1 }
+  &\equiv \sum_{\ell=0}^{m-p-1} { \ell + p - 1 \choose p - 1 }
   \p{m-p-\ell} \p{\ell+m-p+F_{\mu np}}
   \p{2\ell + F_{\mu np}}
-  \prod_{q=1}^{p-1} \p{\ell-q+F_{\mu np}}.
+  \prod_{q=1}^{p-1} \p{\ell-q+F_{\mu np}} \\
+  &= { m \choose p + 1 } \prod_{q=0}^p \p{m-p-q+F_{\mu np}}.
 \end{align}
-
-% \vspace{3cm}
-
-% With reference to \eqref{eq:E_mn_p_G}, we finally have
-% \begin{align}
-%   \sp{S_\mu^m, S_\bmu^n}_-
-%   = C_{\mu mn}
-%   = mn S_\bmu^{n-1} S_\mu^{m-1}
-%   \sum_{q\in\set{0,1}} \p{m-n}^{1-q}\p{2\mu S_z}^q
-%   + \sum_{p=1}^{\min\set{m,n}-1} E_{\mu mn}^{(p)}.
-% \end{align}
-
-% \begin{align}
-%   \prod_{j=1}^{p-1} \p{2k_j+j-n+2\mu S_z}
-%   = \sum_{q=0}^{p-1} \p{2\mu S_z}^q
-%   \sum_{c\in\C_{p-1-q}\p{\Z_{p-1}}} \prod_{j\in c} \p{2k_{j+1}+j+1-n},
-%   \label{eq:product_expansion}
-% \end{align}
-% where $\C\p{k,\S}$ denotes the set of $k$-combinations (i.e.~subsets
-% with $k$ elements) of $\S$; the number of elements in
-% $\C_{p-1-q}\p{\Z_{p-1}}$ is
-% ${ p-1 \choose p-1-q } = { p-1 \choose q }$.
-
-\vspace{3cm}
-
-To derive a system of equations for the two-axis twisting Hamiltonian
-we need to know $\sp{S_\mu^2,S_\bmu^n}_-$, so we expand
+By substitution of $F_{mu np}=p-n+2\mu S_z$, we can expand
 \begin{align}
-  \sp{S_\mu^2,S_\bmu^n}_-
-  = C_{\mu mn}
-  = D_{\mu,2,n} + n D_{\mu,1,n-1} \p{1 - n + 2\mu S_z}
+  \prod_{q=0}^p \p{m-p-q+F_{\mu np}}
+  = \prod_{q=0}^p \p{m-n-q+2\mu S_z}
+  = \sum_{q=0}^{p+1} \p{2\mu S_z}^q e_{p+1-q}\p{m-n-\Z_{p+1}},
+\end{align}
+where $m-n-\Z_{p+1}=\set{m-n-j:j\in\Z_{p+1}}$ and $e_k\p{\S}$ is the
+elementary symmetric polynomial of degree $k$ in elements of $\S$,
+i.e.~the sum of all products of $k$ distinct elements chosen from
+$\S$.  Putting everything together, we have that
+\begin{align}
+  E_{\mu mn}^{(p)}
+  = { m \choose p + 1 } { n \choose p + 1 } \p{p+1}!
+  S_\bmu^{n-p-1} S_\mu^{m-p-1}
+  \sum_{q=0}^{p+1} \p{2\mu}^q e_{p+1-q}\p{m-n-\Z_{p+1}} S_z^q,
+\end{align}
+in terms of which
+\begin{align}
+  \sp{S_\mu^m, S_\bmu^n}_-
+  = mn S_\bmu^{n-1} S_\mu^{m-1} \p{m-n+2\mu S_z}
+  + \sum_{p=1}^{\min\set{m,n}-1} E_{\mu mn}^{(p)}.
 \end{align}
@@
 \newpage
-
 \section{Decoherence of Heisenberg spin operators}
@@
 \end{align}
-Putting everything together, we have that
+Altogether, we have that
 \begin{align}

### 2018-08-21 soc_squeezing/cumulant_expansions.tex
slightly simplified final expression for commutator
fixed typo
added recurrence relation for elementary symmetric polynomial

@@
   \prod_{j=2}^{p-1} \p{2k_{p-j}-j+F_{\mu np}}
-  f_{\mu n\ell}\p{k_{p-2},1} \\
+  f_{\mu np\ell}\p{k_{p-2},1} \\
   &= \f12 \sum_{\p{\v k,p-2,\ell}}
@@
 i.e.~the sum of all products of $k$ distinct elements chosen from
-$\S$.  Putting everything together, we have that
+$\S$.  Among other ways, the elementary symmetric polynomial
+$e_k\p{\S_n}$ for $\S_n\equiv\set{x_1,x_2,\cdots,x_n}$ can be simply
+computed via the recurrence relation
 \begin{align}
-  E_{\mu mn}^{(p)}
-  = { m \choose p + 1 } { n \choose p + 1 } \p{p+1}!
-  S_\bmu^{n-p-1} S_\mu^{m-p-1}
-  \sum_{q=0}^{p+1} \p{2\mu}^q e_{p+1-q}\p{m-n-\Z_{p+1}} S_z^q,
+  e_k\p{\S_n} = e_k\p{\S_{n-1}} + x_n e_{k-1}\p{\S_{n-1}}
 \end{align}
-in terms of which
+together with the terminal conditions $e_k\p{\S_k}=\prod_{j=1}^k x_j$
+and $e_0\p{\S_n}=1$.  Putting everything together, we have
+\begin{align}
+  E_{\mu mn}^{(p-1)}
+  = p! { m \choose p } { n \choose p }
+  S_\bmu^{n-p} S_\mu^{m-p}
+  \sum_{q=0}^p \p{2\mu}^q e_{p-q}\p{m-n-\Z_p} S_z^q,
+\end{align}
+where in the expressed form, $E_{\mu mn}^{(0)} = D_{\mu mn}$, so
+finally
 \begin{align}
   \sp{S_\mu^m, S_\bmu^n}_-
-  = mn S_\bmu^{n-1} S_\mu^{m-1} \p{m-n+2\mu S_z}
-  + \sum_{p=1}^{\min\set{m,n}-1} E_{\mu mn}^{(p)}.
+  = \sum_{p=1}^{\min\set{m,n}} E_{\mu mn}^{(p-1)}.
 \end{align}

### 2018-08-21 soc_squeezing/cumulant_expansions.tex
simplified commutator derivation

@@
   = \f{n!}{\p{n-p}!}
-  \sum_{k_1=0}^{m-2}\cdots\sum_{k_p=k_{p-1}}^{m-p-1}
+  \sum_{k_1=0}^{m-2} \sum_{k_2=k_1}^{m-3} \cdots\sum_{k_p=k_{p-1}}^{m-p-1}
   D_{\mu,m-p,n-p}^{(k_p)} \prod_{j=1}^p \p{2k_j+j-n+2\mu S_z}.
@@
 
-Observing that we always have $k_1\le k_2\le\cdots\le k_p$, we can
-reverse the order of the sums in \eqref{eq:E_mn_p} and relabel
-$k_p\to\ell$ to get
-\begin{align}
-  E_{\mu mn}^{(p)}
-  = \f{n!}{\p{n-p}!}
-  \sum_{\ell=0}^{m-p-1} D_{\mu,m-p,n-p}^{(\ell)} \p{2\ell+p-n+2\mu S_z}
-  \sum_{\substack{\v k\in\Z_{\ell+1}^{p-1}\\
-      k_{p-2}\ge k_{p-1}\ge\cdots\ge k_1}}
-  \prod_{j=1}^{p-1} \p{2k_j+j-n+2\mu S_z},
-\end{align}
-where $\Z_q$ denotes the set of $q$ smallest non-negative integers,
-e.g.~$\Z_{\ell+1}=\set{0,1,\cdots,\ell}$.  Defining for shorthand
-\begin{align}
-  \sum_{\p{\v k,p,\ell}} X \equiv
-  \sum_{\substack{\v k\in\Z_{\ell+1}^{p-1}\\
-      k_{p-1}\ge k_{p-2}\ge\cdots\ge k_1}} X,
-  &&
-  F_{\mu np} \equiv p - n + 2\mu S_z,
-\end{align}
-we can say
+Observing that we always have
+$0\le k_1\le k_2\le\cdots\le k_p\le m-p-1$, we can rearrange the order
+of the sums in \eqref{eq:E_mn_p} and relabel $k_p\to\ell$ to get
 \begin{align}
@@
   \sum_{\ell=0}^{m-p-1} D_{\mu,m-p,n-p}^{(\ell)} \p{2\ell+F_{\mu np}}
-  \sum_{\p{\v k,p,\ell}} \prod_{j=1}^{p-1} \p{2k_{p-j}-j+F_{\mu np}}.
+  \sum_{\p{\v k,p,\ell}} \prod_{j=1}^{p-1} \p{2k_{p-j}-j+F_{\mu np}},
   \label{eq:E_mn_p_sum}
+\end{align}
+where for shorthand we define
+\begin{align}
+  F_{\mu np} \equiv p - n + 2\mu S_z,
+  &&
+  \sum_{\p{\v k,p,\ell}} X \equiv
+  \sum_{k_1=0}^\ell \sum_{k_2=k_1}^\ell
+  \cdots \sum_{k_{p-1}=k_{p-2}}^\ell X.
 \end{align}
@@
 \end{align}
-and evaluate sums successively over $k_{p-1},k_{p-2},\cdots,k_2,k_1$
-respectively from $k_{p-2},k_{p-3},\cdots,k_1,0$ to $\ell$, finding
+and evaluate sums successively over $k_{p-1},k_{p-2},\cdots,k_1$,
+finding
 \begin{align}
   \sum_{\p{\v k,p,\ell}} \prod_{j=1}^{p-1} \p{2k_{p-j}-j+F_{\mu np}}
-  &= \sum_{\p{\v k;p-1;\ell}}
-  \prod_{j=2}^{p-1} \p{2k_{p-j}-j+F_{\mu np}}
-  f_{\mu np\ell}\p{k_{p-2},1} \\
-  &= \f12 \sum_{\p{\v k,p-2,\ell}}
-  \prod_{j=3}^{p-1} \p{2k_{p-j}-j+F_{\mu np}}
-  f_{\mu np\ell}\p{k_{p-3},1} f_{\mu np\ell}\p{k_{p-3},2} \\
   &= \f1{r!} \sum_{\p{\v k,p-r,\ell}}
@@
 Substitution of this result together with $D_{\mu,m-p,n-p}^{(\ell)}$
-as given in using \eqref{eq:D_mn_k} into $E_{\mu mn}^{(p)}$ as
-expressed in \eqref{eq:E_mn_p_sum} then gives us
+with \eqref{eq:D_mn_k} into \eqref{eq:E_mn_p_sum} then gives us
 \begin{align}
@@
   \prod_{q=1}^{p-1} \p{\ell-q+F_{\mu np}} \\
-  &= { m \choose p + 1 } \prod_{q=0}^p \p{m-p-q+F_{\mu np}}.
+  &= { m \choose p + 1 } \prod_{q=0}^p \p{m-p-q+F_{\mu np}} \\
+  &= { m \choose p + 1 } \prod_{q=0}^p \p{m-n-q+2\mu S_z} \\
+  &= { m \choose p + 1 } \sum_{q=0}^{p+1} \p{2\mu S_z}^q
+  e_{p+1-q}\p{m-n-\Z_{p+1}},
 \end{align}
-By substitution of $F_{mu np}=p-n+2\mu S_z$, we can expand
-\begin{align}
-  \prod_{q=0}^p \p{m-p-q+F_{\mu np}}
-  = \prod_{q=0}^p \p{m-n-q+2\mu S_z}
-  = \sum_{q=0}^{p+1} \p{2\mu S_z}^q e_{p+1-q}\p{m-n-\Z_{p+1}},
-\end{align}
-where $m-n-\Z_{p+1}=\set{m-n-j:j\in\Z_{p+1}}$ and $e_k\p{\S}$ is the
+where $\Z_k$ denotes the set of $k$ smallest non-negative integers;
+$m-n-\Z_{p+1}\equiv\set{m-n-j:j\in\Z_{p+1}}$; and $e_k\p{\S}$ is the
 elementary symmetric polynomial of degree $k$ in elements of $\S$,
@@
 $\S$.  Among other ways, the elementary symmetric polynomial
-$e_k\p{\S_n}$ for $\S_n\equiv\set{x_1,x_2,\cdots,x_n}$ can be simply
-computed via the recurrence relation
+$e_k\p{\S_n}$ for $\S_n\equiv\set{x_1,x_2,\cdots,x_n}$ can be computed
+via the recurrence relation
 \begin{align}
@@
 together with the terminal conditions $e_k\p{\S_k}=\prod_{j=1}^k x_j$
-and $e_0\p{\S_n}=1$.  Putting everything together, we have
+and $e_0\p{\S_n}=1$.  Finally, we can say that
 \begin{align}
@@
 \end{align}
-where in the expressed form, $E_{\mu mn}^{(0)} = D_{\mu mn}$, so
-finally
+where in this form $E_{\mu mn}^{(0)} = D_{\mu mn}$, so
 \begin{align}

### 2018-08-22 soc_squeezing/cumulant_expansions.tex
more modifications to the commutator derivation...

@@
 \end{align}
-where the first terms in the sum over $p$ is
+where the first two terms in the sum over $p$ are
 \begin{align}
   E_{\mu mn}^{(1)}
-  = n \sum_{k=0}^{m-2} D_{\mu,m-1,n-1}^{(k)}
-  \p{2k+1-n+2\mu S_z},
-\end{align}
-the second
-\begin{align}
+  &= n \sum_{k=0}^{m-2} D_{\mu,m-1,n-1}^{(k)}
+  \p{2k+1-n+2\mu S_z}, \\
   E_{\mu mn}^{(2)}
-  = n \sum_{k_1=0}^{m-2} \p{n-1} \sum_{k_2=k_1}^{m-3}
+  &= n \sum_{k_1=0}^{m-2} \p{n-1} \sum_{k_2=k_1}^{m-3}
   D_{\mu,m-2,n-2}^{(k_2)}
@@
 In principle, the expressions in \eqref{eq:D_mn} and
-\eqref{eq:C_mn_E}-\eqref{eq:E_mn_p} provide a complete expansion for
-the commutator $\sp{S_\mu^m,S_\nu^n}_- = C_{\mu mn}$, but the result
-is -- to put it lightly -- quite a mess: the expression for
+\eqref{eq:C_mn_E}-\eqref{eq:E_mn_p} can be used to evaluate the
+commutator $\sp{S_\mu^m,S_\nu^n}_- = C_{\mu mn}$, but these result are
+-- to put it lightly -- quite a mess: the expression for
 $E_{\mu mp}^{(p)}$ in \eqref{eq:E_mn_p} involves a sum over $p$
@@
 Observing that we always have
-$0\le k_1\le k_2\le\cdots\le k_p\le m-p-1$, we can rearrange the order
-of the sums in \eqref{eq:E_mn_p} and relabel $k_p\to\ell$ to get
+$0\le k_1\le k_2\le\cdots\le k_p\le m-p-1$ in \eqref{eq:E_mn_p}, we
+can rearrange the order of the sums and relabel $k_p\to\ell$ to get
 \begin{align}
@@
   \sum_{\p{\v k,p,\ell}} \prod_{j=1}^{p-1} \p{2k_{p-j}-j+F_{\mu np}}
+  &= \sum_{\p{\v k,p-1,\ell}}
+  \prod_{j=2}^{p-1} \p{2k_{p-j}-j+F_{\mu np}}
+  f_{\mu np\ell}\p{k_{p-2},1} \\
   &= \f1{r!} \sum_{\p{\v k,p-r,\ell}}
@@
   \prod_{q=1}^{p-1} \p{\ell-q+F_{\mu np}} \\
-  &= { m \choose p + 1 } \prod_{q=0}^p \p{m-p-q+F_{\mu np}} \\
-  &= { m \choose p + 1 } \prod_{q=0}^p \p{m-n-q+2\mu S_z} \\
-  &= { m \choose p + 1 } \sum_{q=0}^{p+1} \p{2\mu S_z}^q
-  e_{p+1-q}\p{m-n-\Z_{p+1}},
+  &= { m \choose p + 1 } \prod_{q=0}^p \p{m-p-q+F_{\mu np}}.
 \end{align}
-where $\Z_k$ denotes the set of $k$ smallest non-negative integers;
+where we can expand
+\begin{align}
+  \prod_{q=0}^p \p{m-p-q+F_{\mu np}}
+  = \sum_{q=0}^{p+1} \p{2\mu S_z}^q
+  = \sum_{q=0}^{p+1} \p{2\mu S_z}^q
+  e_{p+1-q}\p{m-n-\Z_{p+1}}.
+\end{align}
+Here $\Z_k$ denotes the set of $k$ smallest non-negative integers;
 $m-n-\Z_{p+1}\equiv\set{m-n-j:j\in\Z_{p+1}}$; and $e_k\p{\S}$ is the
@@
 i.e.~the sum of all products of $k$ distinct elements chosen from
-$\S$.  Among other ways, the elementary symmetric polynomial
-$e_k\p{\S_n}$ for $\S_n\equiv\set{x_1,x_2,\cdots,x_n}$ can be computed
-via the recurrence relation
+$\S$.  The elementary symmetric polynomial $e_k\p{\S_n}$ for
+$\S_n\equiv\set{x_1,x_2,\cdots,x_n}$ can be computed (among other
+ways) via the recurrence relation
 \begin{align}

### 2018-08-23 soc_squeezing/cumulant_expansions.tex
rearranged appendices and started generalizing the decoherence operator

@@
   \f{d}{dt} \O
-  = i\sp{H,\O}_- + \gamma\D\p{\O},
+  = i\sp{H,\O}_- + \gamma\D_-\p{\O},
   &&
-  \D\p{\O} \equiv \sum_j\p{\sigma_j^+\O\sigma_j^-
+  \D_-\p{\O} \equiv \sum_j\p{\sigma_j^+\O\sigma_j^-
     - \f12\sp{\sigma_j^+\sigma_j^-,\O}_+},
@@
   \f{d}{dt}\p{S_z^n}
-  = \gamma \D\p{S_z^n}
+  = \gamma \D_-\p{S_z^n}
   = \gamma \p{S + S_z} \sum_{k=0}^{n-1}
@@
 \end{align}
-where $S\equiv N/2$ and $\D\p{S_z^n}$ is worked out in Appendix
+where $S\equiv N/2$ and $\D_-\p{S_z^n}$ is worked out in Appendix
 \ref{sec:decoherence}.  In particular,
@@
   = i\chi \sp{S_z^2, S_+^\ell S_-^m S_z^n}_-
-  + \gamma \D\p{S_+^\ell S_-^m S_z^n},
+  + \gamma \D_-\p{S_+^\ell S_-^m S_z^n},
 \end{align}
-where we can use identity \eqref{eq:comm_z_mu_nu_pow} to simplify
+where we can use identity \eqref{eq:comm_z_mu} to simplify
 \begin{align}
@@
   = S_+^\ell S_-^m \p{\sp{\ell-m}^2 + 2\sp{\ell-m} S_z} S_z^n,
+  \label{eq:S_z^2_comm}
 \end{align}
-and the effect of decoherence through $\D\p{S_+^\ell S_-^m S_z^n}$ is
+and the effect of decoherence through $\D_-\p{S_+^\ell S_-^m S_z^n}$ is
 worked out in Appendix \ref{sec:decoherence}.  In order to compute
@@
 
-\section{One-axis twisting with a transverse field}
+% \section{One-axis twisting with a transverse field}
 
-We now consider again squeezing the initial state $\ket{X}$, but this
-time with the transverse field Hamiltonian
+% We now consider again squeezing the initial state $\ket{X}$, but this
+% time with the transverse field Hamiltonian
+% \begin{align}
+%   H_{\t{TVF}}
+%   \equiv \chi S_z^2- \Omega S_x
+%   = \chi S_z^2- \f{\Omega}{2} \p{S_+ + S_-}.
+% \end{align}
+% To determine the coherent evolution of the operator
+% $S_+^\ell S_-^m S_z^n$, we now need to compute
+% \begin{align}
+%   \sp{H_{\t{TVF}}, S_+^\ell S_-^m S_z^n}_-
+%   = \chi \sp{S_z^2, S_+^\ell S_-^m S_z^n}_-
+%   - \f{\Omega}{2} \sp{S_+, S_+^\ell S_-^m S_z^n}_-
+%   - \f{\Omega}{2} \sp{S_-, S_+^\ell S_-^m S_z^n}_-.
+% \end{align}
+% The first commutator above is identical to that for OAT, and is given
+% in \eqref{eq:S_z^2_comm}.  For the second commutator, we expand
+% \begin{align}
+%   \sp{S_+, S_+^\ell S_-^m S_z^n}_-
+%   = S_+^{\ell+1} S_-^m S_z^n - S_+^\ell S_-^m S_z^n S_+,
+% \end{align}
+% where
+% \begin{align}
+%   S_-^m S_z^n S_+
+%   = S_-^m S_+ \p{1 + S_z}^n
+%   = \sp{S_+ S_-^m - m S_-^{m-1} \p{1 - m + 2 S_z}}
+%   \sum_{k=0}^n { n \choose k } S_z^k,
+% \end{align}
+% so
+% \begin{align}
+%   \sp{S_+, S_+^\ell S_-^m S_z^n}_-
+%   = S_+^{\ell+1} S_-^m S_z^n
+%   - S_+^\ell \sp{S_+ S_-^m - m S_-^{m-1} \p{1 - m + 2 S_z}}
+%   \sum_{k=0}^n { n \choose k } S_z^k.
+% \end{align}
+% Similarly,
+% \begin{align}
+%   \sp{S_-, S_+^\ell S_-^m S_z^n}_-
+%   = S_- S_+^\ell S_-^m S_z^n - S_+^\ell S_-^m S_z^n S_-,
+% \end{align}
+% where
+% \begin{align}
+%   S_- S_+^\ell S_-^m
+%   = S_+^\ell S_-^{m+1} + \ell S_+^{\ell-1}\p{1-\ell-2S_z} S_-^m
+%   = S_+^\ell S_-^{m+1} + \ell S_+^{\ell-1} S_-^m \p{1-m-\ell-2S_z},
+% \end{align}
+%   and
+% \begin{align}
+%   S_z^n S_- = S_- \p{1 - S_z}^n
+%   = S_- \sum_{k=0}^n \p{-1}^k { n \choose k } S_z^k,
+% \end{align}
+% so
+% \begin{align}
+%   \sp{S_-, S_+^\ell S_-^m S_z^n}_-
+%   = S_+^\ell S_-^{m+1} S_z^n
+%   + \ell S_+^{\ell-1} S_-^m \p{1-m-\ell-2S_z} S_z^n
+%   - S_+^\ell S_-^{m+1} \sum_{k=0}^n \p{-1}^k { n \choose k } S_z^k.
+% \end{align}
+
+\section{Evolution under general Hamiltonians with decoherence}
+
+All collective spin Hamiltonians can be written in the form
 \begin{align}
-  H_{\t{TVF}}
-  \equiv \chi S_z^2- \Omega S_x.
+  H = \sum_{p,q,r} h_{pqr} S_+^p S_-^q S_z^r,
 \end{align}
-To determine the coherent evolution of the operator
-$S_\mu^m S_\nu^n S_z^n$, we now need to compute
+for which the coherent evolution of a general operator
+$S_+^\ell S_-^m S_z^n$ is given by
 \begin{align}
-  \sp{H_{\t{TVF}}, S_\mu^m S_\nu^n S_z^n}_-
-  = \chi \sp{S_z^2, S_\mu^m S_\nu^n S_z^n}_-
-  - \Omega \sp{S_x, S_\mu^m S_\nu^n S_z^n}_-.
+  \f{d}{dt} \p{S_+^\ell S_-^m S_z^n}
+  = i\sum_{p,q,r} h_{pqr} \sp{S_+^p S_-^q S_z^r, S_+^\ell S_-^m S_z^n}_-
+  + \D_-\p{S_+^\ell S_-^m S_z^n}.
 \end{align}
-Here first commutator is identical to that for OAT, while
 
@@
 The elementary commutation relations between single-weight spin
-operators are
+operators are, with $\bmu\equiv-\mu\in\set{+1,-1}$ for brevity,
 \begin{align}
@@
   &= \mu S_\mu,
-  \label{eq:comm_z_mu} \\
-  \sp{\sigma_j^\mu,\sigma_k^\nu}_-
-  &= \delta_{jk} \delta_{\mu,-\nu} \mu \sigma_j^z,
+  \label{eq:comm_z_mu_base} \\
+  \sp{\sigma_j^\mu,\sigma_k^\bmu}_-
+  &= \delta_{jk} \mu \sigma_j^z,
   &
-  \sp{S_\mu,\sigma_j^\nu}_-
-  &= \sp{\sigma_j^\mu,S_\nu}_-
-  = \delta_{\mu,-\nu} \mu \sigma_j^z,
+  \sp{S_\mu,\sigma_j^\bmu}_-
+  &= \sp{\sigma_j^\mu,S_\bmu}_- = \mu \sigma_j^z,
   &
-  \sp{S_\mu,S_\nu}_-
-  &= \delta_{\mu,-\nu} 2 \mu S_z.
-  \label{eq:comm_mu_nu}
+  \sp{S_\mu,S_\bmu}_-
+  &= 2 \mu S_z.
+  \label{eq:comm_mu_base}
 \end{align}
@@
   = S_\mu^m \p{\mu m + S_z},
-  \label{eq:push_z_mu}
+  \label{eq:push_z_mu_single}
 \end{align}
@@
 \begin{align}
-  S_z S_\mu^m S_\nu^n
-  = S_\mu^m \p{\mu m + S_z} S_\nu^n
-  = S_\mu^m S_\nu^n \p{\mu m + \nu n + S_z}.
-  \label{eq:push_z_mu_nu}
+  S_z S_\mu^m S_\bmu^n
+  = S_\mu^m \p{\mu m + S_z} S_\bmu^n
+  = S_\mu^m S_\bmu^n \p{\mu\sp{m - n} + S_z}.
 \end{align}
@@
 \begin{align}
-  S_z^\ell S_\mu^m S_\nu^n
-  = S_\mu^m S_\nu^n \p{\mu m + \nu n + S_z}^\ell,
+  S_z^\ell S_\mu^m S_\bmu^n
+  = S_\mu^m S_\bmu^n \p{\mu\sp{m - n} + S_z}^\ell,
+  \label{eq:push_z_mu}
 \end{align}
@@
 \begin{align}
-  \sp{S_z^\ell, S_\mu^m S_\nu^n}_-
-  = S_\mu^m S_\nu^n \p{\sp{\mu m + \nu n + S_z}^\ell - S_z^\ell}
-  = S_\mu^m S_\nu^n  \sum_{k=0}^{\ell-1} {\ell \choose k}
-  \p{\mu m + \nu n}^{\ell-k} S_z^k.
-  \label{eq:comm_z_mu_nu_pow}
+  \sp{S_z^\ell, S_\mu^m S_\bmu^n}_-
+  = S_\mu^m S_\bmu^n \sp{\mu^\ell \p{\mu\sp{m - n} + S_z}^\ell - S_z^\ell}
+  = S_\mu^m S_\bmu^n  \sum_{k=0}^{\ell-1} {\ell \choose k}
+  \mu^{\ell-k} \p{m-n}^{\ell-k} S_z^k.
+  \label{eq:comm_z_mu}
 \end{align}
 Similarly finding commutation relations between high-weight transverse
-spin operators, i.e.~$S_\mu^m$ and $S_\nu^n$, turns out to be
+spin operators, i.e.~$S_\mu^m$ and $S_\bmu^n$, turns out to be
 considerably more difficult than the cases we have worked out thus
@@
 
-\newpage
+\section{Decoherence of Heisenberg spin operators}
+\label{sec:decoherence}
+
+The decoherence operator for uncorrelated incoherent decay ($\mu=-1$)
+or excitation ($\mu=1$) of individual spins is, with $\bmu\equiv-\mu$,
+\begin{align}
+  \D_\mu\p{\O}
+  \equiv \sum_j\p{\sigma_j^\bmu \O \sigma_j^\mu
+    - \f12\sp{\sigma_j^\bmu \sigma_j^\mu,\O}_+},
+\end{align}
+where
+\begin{align}
+  \sum_j \f12\sp{\sigma_j^\bmu \sigma_j^\mu,\O}_+
+  = \f12\sp{S-\mu S_z,\O}_+
+  = S\O - \f{\mu}{2}\sp{S_z,\O}_+
+  = \O\p{S-\mu S_z} - \f{\mu}{2}\sp{S_z,\O}_-,
+\end{align}
+so
+\begin{align}
+  \D_\mu\p{\O}
+  = \sum_j \sigma_j^\bmu \O \sigma_j^\mu
+  - \O\p{S - \mu S_z} + \f{\mu}{2} \sp{S_z,\O}_-.
+\end{align}
+We are interested in the effect of this decoherence on the dynamics of
+collective spin operators of the form $S_\bmu^\ell S_\mu^m S_z^n$.  We
+thus expand
+\begin{align}
+  \D_\mu\p{S_\bmu^\ell S_\mu^m S_z^n}
+  = \sum_j \sigma_j^\bmu S_\bmu^\ell S_\mu^m S_z^n \sigma_j^\mu
+  - S_\bmu^\ell S_\mu^m S_z^n \p{S - \mu S_z}
+  + \f{\mu}{2} \sp{S_z, S_\bmu^\ell S_\mu^m S_z^n}_-.
+  \label{eq:D_z+-}
+\end{align}
+The commutator in \eqref{eq:D_z+-} is straightforward to simplify
+using \eqref{eq:comm_z_mu}:
+\begin{align}
+  \mu \sp{S_z, S_\bmu^\ell S_\mu^m S_z^n}_-
+  = \mu \sp{S_z, S_\bmu^\ell S_\mu^m}_- S_z^n
+  = \mu\bmu \p{\ell-m} S_\bmu^\ell S_\mu^m S_z^n
+  = - \p{\ell-m} S_\bmu^\ell S_\mu^m S_z^n,
+\end{align}
+while the first term in \eqref{eq:D_z+-} can be written in terms of
+collective spin operators if we use
+\eqref{eq:comm_z_mu_base}-\eqref{eq:comm_mu_base} to push the jump
+operators together:
+\begin{align}
+  \sum_j \sigma_j^\bmu S_\bmu^\ell S_\mu^m S_z^n \sigma_j^\mu
+  = \sum_j S_\bmu^\ell \sigma_j^\bmu \sigma_j^\mu S_\mu^m \p{\mu+S_z}^n
+  = S_\bmu^\ell \p{S - \mu S_z} S_\mu^m
+  \sum_{k=0}^n \mu^{n-k} { n \choose k } S_z^k.
+\end{align}
+Finally, we need to push the middle $S_z$ to the right, which we can
+do using \eqref{eq:push_z_mu_single}:
+\begin{align}
+  \sum_j \sigma_j^\bmu S_\bmu^\ell S_\mu^m S_z^n \sigma_j^\mu
+  = S_\bmu^\ell S_\mu^m \p{S - m - \mu S_z}
+  \sum_{k=0}^n \mu^{n-k} { n \choose k } S_z^k.
+\end{align}
+Altogether, we have that
+\begin{align}
+  \D_\mu\p{S_\bmu^\ell S_\mu^m S_z^n}
+  &= S_\bmu^\ell S_\mu^m \sp{\p{S - m - \mu S_z} \sum_{k=0}^n
+    \mu^{n-k} { n \choose k } S_z^k
+    - \p{S - \mu S_z + \f12\sp{\ell-m}} S_z^n} \\
+  &= S_\bmu^\ell S_\mu^m \sp{\p{S - m - \mu S_z} \sum_{k=0}^{n-1}
+    \mu^{n-k} { n \choose k } S_z^k - \f12 \p{\ell+m} S_z^n}.
+\end{align}
+
 \section{Commutators between high-weight transverse spin operators}
@@
 transverse to the decoherence (i.e.~$z$) axis, we compute (with
-$\bmu\equiv-\mu$ for brevity)
+$\bmu\equiv-\mu\in\set{+1,-1}$)
 \begin{align}
   S_\mu S_\bmu^n
-  &= \p{2\mu S_z + S_\bmu S_\mu} S_\bmu^{n-1} \\
-  &= 2\mu S_z S_\bmu^{n-1} + S_\bmu S_\mu S_\bmu^{n-1} \\
-  &= 2\mu S_z S_\bmu^{n-1}
-  + S_\bmu \sp{2\mu S_z S_\bmu^{n-2}
-    + S_\bmu S_\mu S_\bmu^{n-2}} \\
-  &= 2\mu \p{S_z S_\bmu^{n-1} + S_\bmu S_z S_\bmu ^{n-2}}
-  + S_\bmu^2 S_\mu S_\bmu^{n-2} \\
-  &= 2\mu \p{S_z S_\bmu^{n-1} + S_\bmu S_z S_\bmu ^{n-2}
-    + S_\bmu^2 S_z S_\bmu ^{n-3}} + S_\bmu^3 S_\mu S_\bmu^{n-3} \\
-  &= 2\mu \sum_{k=0}^{n-1} S_\bmu^k S_z S_\bmu^{n-k-1}
-  + S_\bmu^n S_\mu.
-  \label{eq:comm_mu_nu_single}
+  &= S_\bmu S_\mu S_\bmu^{n-1} + 2\mu S_z S_\bmu^{n-1} \\
+  &= S_\bmu \sp{S_\bmu S_\mu S_\bmu^{n-2} + 2\mu S_z S_\bmu^{n-2}}
+  + 2\mu S_z S_\bmu^{n-1} \\
+  &= S_\bmu^2 S_\mu S_\bmu^{n-2}
+  + 2\mu \p{S_\bmu S_z S_\bmu ^{n-2} + S_z S_\bmu^{n-1}} \\
+  &= S_\bmu^n S_\mu + 2\mu \sum_{k=0}^{n-1} S_\bmu^k S_z S_\bmu^{n-k-1}.
+  \label{eq:comm_mu_single}
 \end{align}
-While \eqref{eq:comm_mu_nu_single} gives us the commutator
+While \eqref{eq:comm_mu_single} gives us the commutator
 $\sp{S_\mu,S_\bmu^n}_-$, we would like to enforce an ordering on
@@
   S_\mu^m S_\bmu^n
-  &= S_\mu^{m-1} \sp{n S_\bmu^{n-1}\p{1 - n + 2\mu S_z}
-    + S_\bmu^n S_\mu} \\
-  &= n \sp{S_\mu^{m-1} S_\bmu^{n-1} \p{1 - n + 2\mu S_z}}
-  + S_\mu^{m-1} S_\bmu^n S_\mu \\
-  &= n \sum_{k=0}^{m-1} S_\mu^{m-k-1} S_\bmu^{n-1}
-  \p{1 - n + 2\mu S_z} S_\mu^k + S_\bmu^n S_\mu^m,
+  &= S_\mu^{m-1} S_\bmu^n S_\mu
+  + n \sp{S_\mu^{m-1} S_\bmu^{n-1} \p{1 - n + 2\mu S_z}} \\
+  &= S_\bmu^n S_\mu^m
+  + n \sum_{k=0}^{m-1} S_\mu^{m-k-1} S_\bmu^{n-1}
+  \p{1 - n + 2\mu S_z} S_\mu^k,
 \end{align}
@@
   \sp{S_\mu^m, S_\bmu^n}_-
-  = C_{\mu mn}
+  = C_{mn;\mu}
   \equiv n \sum_{k=0}^{m-1} S_\mu^{m-k-1} S_\bmu^{n-1}
@@
 \end{align}
-We now need rearrange the operators in $C_{\mu mn}$ into a standard
+We now need rearrange the operators in $C_{mn;\mu}$ into a standard
 order, which means pushing all $S_z$ operators to the right and, for
@@
 \begin{align}
-  C_{\mu mn}
+  C_{mn;\mu}
   &= n \sum_{k=0}^{m-1}
-  \p{S_\bmu^{n-1} S_\mu^{m-k-1} + C_{\mu,m-k-1,n-1}} S_\mu^k
+  \p{S_\bmu^{n-1} S_\mu^{m-k-1} + C_{m-k-1,n-1;\mu}} S_\mu^k
   \p{2k + 1 - n + 2\mu S_z} \\
-  &= D_{\mu mn}
-  + n \sum_{k=0}^{m-2} C_{\mu,m-k-1,n-1}
+  &= D_{mn;\mu}
+  + n \sum_{k=0}^{m-2} C_{m-k-1,n-1;\mu}
   S_\mu^k \p{2k + 1 - n + 2\mu S_z},
@@
 where we have dropped the last term in the final sum because
-$C_{\mu,m-k-1,n-1}=\sp{S_\mu^{m-k-1},S_\bmu^{n-1}}_-=0$ if $k=m-1$,
-and
+$C_{m-k-1,n-1;\mu}=0$ if $k=m-1$, and
 \begin{align}
-  D_{\mu mn}
+  D_{mn;\mu}
   \equiv mn S_\bmu^{n-1} S_\mu^{m-1} \p{m - n + 2\mu S_z}.
@@
 \end{align}
-To our despair, we have arrived in \eqref{eq:C_mn} to a {\it
-  recursive} expression for $C_{\mu mn}$.  Furthermore, we have not
-even managed to order the spin operators, as $C_{\mu,m-k-1,n-1}$
-contains $S_z$ operators which are to the left of $S_\mu^k$.  To sort
-all spin operators once and for all, we define
+To our despair, we have arrived in \eqref{eq:C_mn} to a recursive
+formula for $C_{mn;\mu}$.  Furthermore, we have not even managed to
+order the spin operators, as $C_{m-k-1,n-1;\mu}$ contains $S_z$
+operators which are to the left of $S_\mu^k$.  To sort all spin
+operators once and for all, we define
 \begin{align}
-  C_{\mu mn}^{(k)} \equiv C_{\mu,m-k,n} S_\mu^k,
+  C_{mn;\mu}^{(k)} \equiv C_{m-k,n;\mu} S_\mu^k,
   &&
-  D_{\mu mn}^{(k)} \equiv D_{\mu,m-k,n} S_\mu^k,
+  D_{mn;\mu}^{(k)} \equiv D_{m-k,n;\mu} S_\mu^k,
 \end{align}
@@
 \begin{align}
-  D_{\mu mn}^{(k)}
-  &\equiv D_{\mu,m-k,n}S_\mu^k \\
+  D_{mn;\mu}^{(k)}
+  &\equiv D_{m-k,n;\mu}S_\mu^k \\
   &= \p{m-k}n S_\bmu^{n-1} S_\mu^{m-k-1}
@@
 \begin{align}
-  C_{\mu mn}^{(k)}
-  &= D_{\mu,m-k,n} S_\mu^k + n \sum_{j=0}^{m-k-2}
-  C_{\mu,m-k-j-1,n-1} S_\mu^j \p{2j+1-n+2\mu S_z} S_\mu^k \\
-  &= D_{\mu mn}^{(k)} + n \sum_{j=0}^{m-k-2}
-  C_{\mu,m-k-j-1,n-1} S_\mu^{j+k} \p{2j+2k+1-n+2\mu S_z} \\
-  &= D_{\mu mn}^{(k)} + n \sum_{j=0}^{m-k-2}
-  C_{\mu,m-1,n-1}^{(k+j)} \p{2\sp{j+k}+1-n+2\mu S_z} \\
-  &= D_{\mu mn}^{(k)} + n \sum_{j=k}^{m-2}
-  C_{\mu,m-1,n-1}^{(j)} \p{2j+1-n+2\mu S_z}.
+  C_{mn;\mu}^{(k)}
+  &= D_{m-k,n;\mu} S_\mu^k + n \sum_{j=0}^{m-k-2}
+  C_{m-k-j-1,n-1;\mu} S_\mu^j \p{2j+1-n+2\mu S_z} S_\mu^k \\
+  &= D_{mn;\mu}^{(k)} + n \sum_{j=0}^{m-k-2}
+  C_{m-k-j-1,n-1;\mu} S_\mu^{j+k} \p{2j+2k+1-n+2\mu S_z} \\
+  &= D_{mn;\mu}^{(k)} + n \sum_{j=0}^{m-k-2}
+  C_{m-1,n-1;\mu}^{(k+j)} \p{2\sp{j+k}+1-n+2\mu S_z} \\
+  &= D_{mn;\mu}^{(k)} + n \sum_{j=k}^{m-2}
+  C_{m-1,n-1;\mu}^{(j)} \p{2j+1-n+2\mu S_z}.
   \label{eq:C_mn_k}
@@
 operators in \eqref{eq:C_mn_k} have been sorted into a standard order.
-We can now repeatedly substitute $C_{\mu mn}^{(k)}$ into itself, each
+We can now repeatedly substitute $C_{mn;\mu}^{(k)}$ into itself, each
 time decreasing $m$ and $n$ by 1, until one of $m$ or $n$ reaches
@@
 \begin{align}
-  C_{\mu mn}
-  = C_{\mu mn}^{(0)}
-  = D_{\mu mn}
-  + \sum_{p=1}^{\min\set{m,n}-1} E_{\mu mn}^{(p)},
+  C_{mn;\mu}
+  = C_{mn;\mu}^{(0)}
+  = D_{mn;\mu}
+  + \sum_{p=1}^{\min\set{m,n}-1} E_{mn;\mu}^{(p)},
   \label{eq:C_mn_E}
@@
 \begin{align}
-  E_{\mu mn}^{(1)}
-  &= n \sum_{k=0}^{m-2} D_{\mu,m-1,n-1}^{(k)}
+  E_{mn;\mu}^{(1)}
+  &= n \sum_{k=0}^{m-2} D_{m-1,n-1;\mu}^{(k)}
   \p{2k+1-n+2\mu S_z}, \\
-  E_{\mu mn}^{(2)}
+  E_{mn;\mu}^{(2)}
   &= n \sum_{k_1=0}^{m-2} \p{n-1} \sum_{k_2=k_1}^{m-3}
-  D_{\mu,m-2,n-2}^{(k_2)}
+  D_{m-2,n-2;\mu}^{(k_2)}
   \p{2k_2+2-n+2\mu S_z} \p{2k_1+1-n+2\mu S_z},
@@
 \begin{align}
-  E_{\mu mn}^{(p)}
+  E_{mn;\mu}^{(p)}
   = \f{n!}{\p{n-p}!}
   \sum_{k_1=0}^{m-2} \sum_{k_2=k_1}^{m-3} \cdots\sum_{k_p=k_{p-1}}^{m-p-1}
-  D_{\mu,m-p,n-p}^{(k_p)} \prod_{j=1}^p \p{2k_j+j-n+2\mu S_z}.
+  D_{m-p,n-p;\mu}^{(k_p)} \prod_{j=1}^p \p{2k_j+j-n+2\mu S_z}.
   \label{eq:E_mn_p}
@@
 \eqref{eq:C_mn_E}-\eqref{eq:E_mn_p} can be used to evaluate the
-commutator $\sp{S_\mu^m,S_\nu^n}_- = C_{\mu mn}$, but these result are
--- to put it lightly -- quite a mess: the expression for
-$E_{\mu mp}^{(p)}$ in \eqref{eq:E_mn_p} involves a sum over $p$
+commutator $\sp{S_\mu^m,S_\bmu^n}_- = C_{mn;\mu}$, but these result
+are -- to put it lightly -- quite a mess: the expression for
+$E_{mn;\mu}^{(p)}$ in \eqref{eq:E_mn_p} involves a sum over $p$
 mutually dependent intermediate variables!  We therefore devote the
 rest of this section to simplifying our result for the commutator
-$\sp{S_\mu^m,S_\nu^n}_-$.
+$\sp{S_\mu^m,S_\bmu^n}_-$.
 
@@
 \begin{align}
-  E_{\mu mn}^{(p)}
+  E_{mn;\mu}^{(p)}
   = \f{n!}{\p{n-p}!}
-  \sum_{\ell=0}^{m-p-1} D_{\mu,m-p,n-p}^{(\ell)} \p{2\ell+F_{\mu np}}
-  \sum_{\p{\v k,p,\ell}} \prod_{j=1}^{p-1} \p{2k_{p-j}-j+F_{\mu np}},
+  \sum_{\ell=0}^{m-p-1} D_{m-p,n-p;\mu}^{(\ell)} \p{2\ell+F_{np;\mu}}
+  \sum_{\p{\v k,p,\ell}} \prod_{j=1}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}},
   \label{eq:E_mn_p_sum}
@@
 \begin{align}
-  F_{\mu np} \equiv p - n + 2\mu S_z,
+  F_{np;\mu} \equiv p - n + 2\mu S_z,
   &&
@@
 \begin{align}
-  f_{\mu np\ell}\p{k,q} \equiv \p{\ell-k+q} \p{\ell+k-q+F_{\mu np}},
+  f_{np\ell;\mu}\p{k,q} \equiv \p{\ell-k+q} \p{\ell+k-q+F_{np;\mu}},
 \end{align}
@@
 \begin{align}
-  \sum_{\p{\v k,p,\ell}} \prod_{j=1}^{p-1} \p{2k_{p-j}-j+F_{\mu np}}
+  \sum_{\p{\v k,p,\ell}} \prod_{j=1}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}}
   &= \sum_{\p{\v k,p-1,\ell}}
-  \prod_{j=2}^{p-1} \p{2k_{p-j}-j+F_{\mu np}}
-  f_{\mu np\ell}\p{k_{p-2},1} \\
+  \prod_{j=2}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}}
+  f_{np\ell;\mu}\p{k_{p-2},1} \\
   &= \f1{r!} \sum_{\p{\v k,p-r,\ell}}
-  \prod_{j=1+r}^{p-1} \p{2k_{p-j}-j+F_{\mu np}}
-  \prod_{q=1}^r f_{\mu np\ell}\p{k_{p-1-r},q} \\
-  &= \f1{\p{p-1}!} \prod_{q=1}^{p-1} f_{\mu np\ell}\p{0,q} \\
+  \prod_{j=1+r}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}}
+  \prod_{q=1}^r f_{np\ell;\mu}\p{k_{p-1-r},q} \\
+  &= \f1{\p{p-1}!} \prod_{q=1}^{p-1} f_{np\ell;\mu}\p{0,q} \\
   &= { \ell + p - 1 \choose p - 1 }
-  \prod_{q=1}^{p-1} \p{\ell-q+F_{\mu np}}.
+  \prod_{q=1}^{p-1} \p{\ell-q+F_{np;\mu}}.
 \end{align}
-Substitution of this result together with $D_{\mu,m-p,n-p}^{(\ell)}$
+Substitution of this result together with $D_{m-p,n-p;\mu}^{(\ell)}$
 with \eqref{eq:D_mn_k} into \eqref{eq:E_mn_p_sum} then gives us
 \begin{align}
-  E_{\mu mn}^{(p)}
-  = \f{n!}{\p{n-p-1}!} S_\bmu^{n-p-1} S_\mu^{m-p-1} G_{\mu mnp}
+  E_{mn;\mu}^{(p)}
+  = \f{n!}{\p{n-p-1}!} S_\bmu^{n-p-1} S_\mu^{m-p-1} G_{mnp;\mu}
 \end{align}
@@
 \begin{align}
-  G_{\mu mnp}
+  G_{mnp;\mu}
   &\equiv \sum_{\ell=0}^{m-p-1} { \ell + p - 1 \choose p - 1 }
-  \p{m-p-\ell} \p{\ell+m-p+F_{\mu np}}
-  \p{2\ell + F_{\mu np}}
-  \prod_{q=1}^{p-1} \p{\ell-q+F_{\mu np}} \\
-  &= { m \choose p + 1 } \prod_{q=0}^p \p{m-p-q+F_{\mu np}}.
+  \p{m-p-\ell} \p{\ell+m-p+F_{np;\mu}}
+  \p{2\ell + F_{np;\mu}}
+  \prod_{q=1}^{p-1} \p{\ell-q+F_{np;\mu}} \\
+  &= { m \choose p + 1 } \prod_{q=0}^p \p{m-p-q+F_{np;\mu}}.
 \end{align}
-where we can expand
+We can further expand
 \begin{align}
-  \prod_{q=0}^p \p{m-p-q+F_{\mu np}}
-  = \sum_{q=0}^{p+1} \p{2\mu S_z}^q
-  = \sum_{q=0}^{p+1} \p{2\mu S_z}^q
-  e_{p+1-q}\p{m-n-\Z_{p+1}}.
+  \prod_{q=0}^p \p{m-p-q+F_{np;\mu}}
+  = \prod_{q=0}^p \p{m-n-q+2\mu S_z}
+  = \sum_{q=0}^{p+1} \p{-1}^{p+1-q} { p+1 \brack q } \p{m-n+2\mu S_z}^q,
 \end{align}
-Here $\Z_k$ denotes the set of $k$ smallest non-negative integers;
-$m-n-\Z_{p+1}\equiv\set{m-n-j:j\in\Z_{p+1}}$; and $e_k\p{\S}$ is the
-elementary symmetric polynomial of degree $k$ in elements of $\S$,
-i.e.~the sum of all products of $k$ distinct elements chosen from
-$\S$.  The elementary symmetric polynomial $e_k\p{\S_n}$ for
-$\S_n\equiv\set{x_1,x_2,\cdots,x_n}$ can be computed (among other
-ways) via the recurrence relation
+where ${ n \brack k }$ is an unsigned Stirling number of the first
+kind, and finally
 \begin{align}
-  e_k\p{\S_n} = e_k\p{\S_{n-1}} + x_n e_{k-1}\p{\S_{n-1}}
+  \sum_{q=0}^p \p{-1}^{p-q} { p \brack q } \p{m-n+2\mu S_z}^q
+  &= \sum_{q=0}^p \p{-1}^{p-q} { p \brack q } \sum_{\ell=0}^q
+  { q \choose \ell } \p{2\mu S_z}^\ell \p{m-n}^{q-\ell} \\
+  &= \sum_{\ell=0}^p \p{2\mu}^\ell S_z^\ell
+  \sum_{q=\ell}^p \p{-1}^{p-q}
+  { p \brack q } { q \choose \ell } \p{m-n}^{q-\ell}.
 \end{align}
-together with the terminal conditions $e_k\p{\S_k}=\prod_{j=1}^k x_j$
-and $e_0\p{\S_n}=1$.  Finally, we can say that
+Putting everything together, we finally have
 \begin{align}
-  E_{\mu mn}^{(p-1)}
+  E_{mn;\mu}^{(p-1)}
   = p! { m \choose p } { n \choose p }
   S_\bmu^{n-p} S_\mu^{m-p}
-  \sum_{q=0}^p \p{2\mu}^q e_{p-q}\p{m-n-\Z_p} S_z^q,
+  \sum_{\ell=0}^p \p{2\mu}^\ell \epsilon_{mnp}^{(\ell)} S_z^\ell,
 \end{align}
-where in this form $E_{\mu mn}^{(0)} = D_{\mu mn}$, so
+with
+\begin{align}
+  \epsilon_{mnp}^{(\ell)}
+  \equiv \sum_{q=\ell}^p \p{-1}^{p-q}
+  { p \brack q } { q \choose \ell } \p{m-n}^{q-\ell},
+  \label{eq:epsilon}
+\end{align}
+where in this final form $E_{mn;\mu}^{(0)} = D_{mn;\mu}$, which
+together with \eqref{eq:C_mn_E} implies that
 \begin{align}
   \sp{S_\mu^m, S_\bmu^n}_-
-  = \sum_{p=1}^{\min\set{m,n}} E_{\mu mn}^{(p-1)}.
+  = \sum_{p=1}^{\min\set{m,n}}
+  p! { m \choose p } { n \choose p } S_\bmu^{n-p} S_\mu^{m-p}
+  \sum_{\ell=0}^p \p{2\mu}^\ell \epsilon_{mnp}^{(\ell)} S_z^\ell.
+  \label{eq:comm_mu}
 \end{align}
 
-\newpage
-\section{Decoherence of Heisenberg spin operators}
-\label{sec:decoherence}
+\section{Commutator between general ordered collective spin operators}
+\label{sec:comm_general}
 
-The decoherence operator for uncorrelated single-spin decay at a
-uniform rate is
+If all collective spin operators are written as ordered products of
+$S_+$, $S_-$, and $S_z$ with all $S_z$ on the right, then the most
+general commutator we need to consider is (with
+$\bmu\equiv-\mu\in\set{+1,-1}$)
 \begin{align}
-  \D\p{\O}
-  \equiv \sum_j\p{\sigma_j^+\O\sigma_j^-
-    - \f12\sp{\sigma_j^+\sigma_j^-,\O}_+},
+  \sp{S_\mu^p S_\bmu^q S_z^a, S_\mu^r S_\bmu^s S_z^b}_-
+  = \S^{pqa}_{rsb;\mu} - \S^{rsb}_{pqa;\mu},
+  &&
+  \S^{pqa}_{rsb;\mu}
+  \equiv S_\mu^p S_\bmu^q S_z^a S_\mu^r S_\bmu^s S_z^b.
 \end{align}
-where
+We can push the $S_z^a$ in $\S^{pqa}_{rsb;\mu}$ to the right through
+$S_\mu^r S_\bmu^s$ to get
 \begin{align}
-  \sum_j \f12\sp{\sigma_j^+\sigma_j^-,\O}_+
-  = \f12\sp{S+S_z,\O}_+
-  = S\O + \f12\sp{S_z,\O}_+
-  = \O\p{S+S_z} + \f12\sp{S_z,\O}_-,
+  \S^{pqa}_{rsb;\mu}
+  = S_\mu^p S_\bmu^q S_\mu^r S_\bmu^s  \p{\mu\sp{r-s}+S_z}^a S_z^b,
 \end{align}
-so
+then commute $S_\bmu^q,S_\mu^r$ to expand
 \begin{align}
-  \D\p{\O}
-  = \sum_j \sigma_j^+\O\sigma_j^- - \O\p{S + S_z} - \f12 \sp{S_z,\O}_-.
+  S_\mu^p S_\bmu^q S_\mu^r S_\bmu^s
+  = S_\mu^{p+r} S_\bmu^{q+s} + \sum_{k=1}^{\min\set{q,r}}
+  k! { q \choose k } { r \choose k } S_\mu^{p+r-k} S_\bmu^{q-k}
+  \sum_{\ell=0}^k \p{-2\mu}^\ell \epsilon_{qrk}^{(\ell)}
+  S_z^\ell S_\bmu^s,
 \end{align}
-We are interested in the effect of this decoherence on the dynamics of
-collective spin operators, all of which are determined by operators of
-the form $S_+^\ell S_-^m S_z^n$.  We thus expand
+and finally push $S_z^\ell$ to the right through $S_\bmu^s$, finding
 \begin{align}
-  \D\p{S_+^\ell S_-^m S_z^n}
-  = \sum_j \sigma_j^+ S_+^\ell S_-^m S_z^n \sigma_j^-
-  - S_+^\ell S_-^m S_z^n \p{S + S_z}
-  - \f12 \sp{S_z, S_+^\ell S_-^m S_z^n}_-.
-  \label{eq:D_z+-}
+  \S^{pqa}_{rsb;\mu}
+  = \sum_{k=0}^{\min\set{q,r}} k! { q \choose k } { r \choose k }
+  S_\mu^{p+r-k} S_\bmu^{q+s-k}
+  \sum_{\ell=0}^k \p{-2\mu}^\ell \epsilon_{qrk}^{(\ell)}
+  Z_{abrs;\mu}^{(\ell)},
 \end{align}
-The last term in \eqref{eq:D_z+-} is straightforward to simplify using
-\eqref{eq:push_z_mu_nu}:
+where $\epsilon_{qrk}^{(\ell)}$ is defined in \eqref{eq:epsilon} and
 \begin{align}
-  \f12 \sp{S_z, S_+^\ell S_-^m S_z^n}_-
-  = \f12 \sp{S_z, S_+^\ell S_-^m}_- S_z^n
-  = \f12 \p{\ell-m} S_+^\ell S_-^m S_z^n,
-\end{align}
-while the first term in \eqref{eq:D_z+-} can be written in terms of
-collective spin operators if we use
-\eqref{eq:comm_z_mu}-\eqref{eq:comm_mu_nu} to push the jump operators
-together:
-\begin{align}
-  \sum_j \sigma_j^+ S_+^\ell S_-^m S_z^n \sigma_j^-
-  = \sum_j S_+^\ell \sigma_j^+ \sigma_j^- S_-^m \p{S_z-1}^n
-  = S_+^\ell \p{S + S_z} S_-^m
-  \sum_{k=0}^n \p{-1}^{n-k} { n \choose k } S_z^k.
-\end{align}
-Finally, we need to push the middle $S_z$ to the right, which we can
-do using \eqref{eq:push_z_mu}:
-\begin{align}
-  \sum_j \sigma_j^+ S_+^\ell S_-^m S_z^n \sigma_j^-
-  = S_+^\ell S_-^m \p{S - m + S_z}
-  \sum_{k=0}^n \p{-1}^{n-k} { n \choose k } S_z^k.
-\end{align}
-Altogether, we have that
-\begin{align}
-  \D\p{S_+^\ell S_-^m S_z^n}
-  &= S_+^\ell S_-^m \sp{\p{S - m + S_z} \sum_{k=0}^n
-    \p{-1}^{n-k} { n \choose k } S_z^k
-    - \p{S + S_z + \f12\sp{\ell-m}} S_z^n} \\
-  &= S_+^\ell S_-^m \sp{\p{S - m + S_z} \sum_{k=0}^{n-1}
-    \p{-1}^{n-k} { n \choose k } S_z^k - \f12 \p{\ell+m} S_z^n}.
+  Z_{abrs;\mu}^{(\ell)}
+  \equiv \p{-\mu s+S_z}^\ell \p{\mu\sp{r-s}+S_z}^a S_z^b
+  = \sum_{m=0}^\ell { \ell \choose m } \p{-\mu s}^{\ell-m}
+  \sum_{n=0}^a { a \choose n } \p{\mu\sp{r-s}}^{a-n} S_z^{b+m+n}.
 \end{align}

### 2018-08-23 soc_squeezing/cumulant_expansions.tex
started writing about TVF and TAT Hamiltonians provided organization for TVF operator vector

@@
 \newcommand{\D}{\mathcal{D}}
+\newcommand{\I}{\mathcal{I}}
 \renewcommand{\O}{\mathcal{O}}
@@
   &&
-  H_{\t{TAT}} = \chi \p{S_z^2 - S_y^2},
+  H_{\t{TAT}}^{z,y} = \chi \p{S_z^2 - S_y^2},
   &&
@@
     \p{-1}^{n-k} { n \choose k } S_z^k - \f12 S_z^n}
+  \label{eq:evolution_OAT_single}
 \end{align}
 and
-\begin{multline}
+\begin{align}
   \f{d}{dt} \p{S_\mu S_\nu S_z^n}
-  = i\chi S_\mu S_\nu \p{\sp{\mu+\nu}^2 + 2 \sp{\mu+\nu} S_z} S_z^n \\
-  + \gamma S_\mu S_\nu
+  &= i\chi S_\mu S_\nu \p{\sp{\mu+\nu}^2 + 2 \sp{\mu+\nu} S_z} S_z^n
+  \tag*{} \\
+  &\quad + \gamma S_\mu S_\nu
   \sp{\p{S - \delta_{\mu,-1} - \delta_{\nu,-1} + S_z} \sum_{k=0}^{n-1}
     \p{-1}^{n-k} { n \choose k } S_z^k - S_z^n},
-\end{multline}
-for $\mu\ge\nu$.  The initial conditions for these systems of equations can be determined from the expansions
+  \eqref{eq:evolution_OAT_pair}
+\end{align}
+for $\mu\ge\nu$.  The initial conditions for these systems of
+equations can be determined from the expansions
 \begin{align}
@@
 
-% \section{One-axis twisting with a transverse field}
+\subsection{Operator imbalance, conservation laws, and time evolution}
 
-% We now consider again squeezing the initial state $\ket{X}$, but this
-% time with the transverse field Hamiltonian
-% \begin{align}
-%   H_{\t{TVF}}
-%   \equiv \chi S_z^2- \Omega S_x
-%   = \chi S_z^2- \f{\Omega}{2} \p{S_+ + S_-}.
-% \end{align}
-% To determine the coherent evolution of the operator
-% $S_+^\ell S_-^m S_z^n$, we now need to compute
-% \begin{align}
-%   \sp{H_{\t{TVF}}, S_+^\ell S_-^m S_z^n}_-
-%   = \chi \sp{S_z^2, S_+^\ell S_-^m S_z^n}_-
-%   - \f{\Omega}{2} \sp{S_+, S_+^\ell S_-^m S_z^n}_-
-%   - \f{\Omega}{2} \sp{S_-, S_+^\ell S_-^m S_z^n}_-.
-% \end{align}
-% The first commutator above is identical to that for OAT, and is given
-% in \eqref{eq:S_z^2_comm}.  For the second commutator, we expand
-% \begin{align}
-%   \sp{S_+, S_+^\ell S_-^m S_z^n}_-
-%   = S_+^{\ell+1} S_-^m S_z^n - S_+^\ell S_-^m S_z^n S_+,
-% \end{align}
-% where
-% \begin{align}
-%   S_-^m S_z^n S_+
-%   = S_-^m S_+ \p{1 + S_z}^n
-%   = \sp{S_+ S_-^m - m S_-^{m-1} \p{1 - m + 2 S_z}}
-%   \sum_{k=0}^n { n \choose k } S_z^k,
-% \end{align}
-% so
-% \begin{align}
-%   \sp{S_+, S_+^\ell S_-^m S_z^n}_-
-%   = S_+^{\ell+1} S_-^m S_z^n
-%   - S_+^\ell \sp{S_+ S_-^m - m S_-^{m-1} \p{1 - m + 2 S_z}}
-%   \sum_{k=0}^n { n \choose k } S_z^k.
-% \end{align}
-% Similarly,
-% \begin{align}
-%   \sp{S_-, S_+^\ell S_-^m S_z^n}_-
-%   = S_- S_+^\ell S_-^m S_z^n - S_+^\ell S_-^m S_z^n S_-,
-% \end{align}
-% where
-% \begin{align}
-%   S_- S_+^\ell S_-^m
-%   = S_+^\ell S_-^{m+1} + \ell S_+^{\ell-1}\p{1-\ell-2S_z} S_-^m
-%   = S_+^\ell S_-^{m+1} + \ell S_+^{\ell-1} S_-^m \p{1-m-\ell-2S_z},
-% \end{align}
-%   and
-% \begin{align}
-%   S_z^n S_- = S_- \p{1 - S_z}^n
-%   = S_- \sum_{k=0}^n \p{-1}^k { n \choose k } S_z^k,
-% \end{align}
-% so
-% \begin{align}
-%   \sp{S_-, S_+^\ell S_-^m S_z^n}_-
-%   = S_+^\ell S_-^{m+1} S_z^n
-%   + \ell S_+^{\ell-1} S_-^m \p{1-m-\ell-2S_z} S_z^n
-%   - S_+^\ell S_-^{m+1} \sum_{k=0}^n \p{-1}^k { n \choose k } S_z^k.
-% \end{align}
+For convenience of language, for collective spin operators of the form
+$S_+^\ell S_-^m S_z^n$ we respectively define the {\it upper}, {\it
+  lower}, and {\it net imbalance}
+\begin{align}
+  \I_+\p{S_+^\ell S_-^m S_z^n} \equiv \ell,
+  &&
+  \I_-\p{S_+^\ell S_-^m S_z^n} \equiv m,
+  &&
+  \I\p{S_+^\ell S_-^m S_z^n} \equiv \ell - m.
+\end{align}
+In the case of one-axis twisting, both upper and lower (and net)
+imbalances are {\it conserved}, meaning that operators with upper and
+lower imbalances $\p{\ell,m}$ depend only on other operators with the
+same imbalances.  We can therefore divide the task of operator time
+evolution into sectors of constant imbalances, which greatly reduces
+computational overhead.  Specifically, operator time evolution now
+involves evolving ``operator vectors''
+\begin{align}
+  \v V_{\t{OAT}}^{(\ell,m)}
+  \equiv S_+^\ell S_-^m \sum_{n\ge0} S_z^n \uv x_n,
+\end{align}
+which in practice must be truncated at some maximal $n=M$.  In the
+case of OAT, the time derivative of $S_+^\ell S_-^m S_z^n$ depends on
+$S_+^\ell S_-^m S_z^{n+1}$ which implies that we can only trust the
+value of $S_+^\ell S_-^m S_z^n$ up to the time at which the initial
+value of $S_+^\ell S_-^m S_z^M$ has (conservatively) an $\O\p{1}$
+contribution to $S_+^\ell S_-^m S_z^n$.
 
-\section{Evolution under general Hamiltonians with decoherence}
+In order to compute squeezing from OAT, we need only need to evolve
+$\v V_{\t{OAT}}^{(1,0)}$, $\v V_{\t{OAT}}^{(0,1)}$,
+$\v V_{\t{OAT}}^{(1,1)}$, and $\v V_{\t{OAT}}^{(2,0)}$.  The values of
+axial spin operators $S_z$ and $S_z^2$ are known exactly at all times
+from \eqref{eq:Sz_OAT}.
+
+\section{Transverse-field and two-axis twisting}
 
@@
 \begin{align}
-  H = \sum_{p,q,r} h_{pqr} S_+^p S_-^q S_z^r,
+  H = \sum_{p,q,r} h_{pqr} S_+^p S_-^q S_z^r.
+  \label{eq:H_general}
 \end{align}
-for which the coherent evolution of a general operator
-$S_+^\ell S_-^m S_z^n$ is given by
+The evolution of a general operator $S_+^\ell S_-^m S_z^n$ under a
+Hamiltonian of the form in \eqref{eq:H_general} is given by
 \begin{align}
@@
   = i\sum_{p,q,r} h_{pqr} \sp{S_+^p S_-^q S_z^r, S_+^\ell S_-^m S_z^n}_-
-  + \D_-\p{S_+^\ell S_-^m S_z^n}.
+  + \gamma \D_-\p{S_+^\ell S_-^m S_z^n},
 \end{align}
+where the commutator is given in Appendix \ref{sec:comm_general}, and
+the decoherence operator in Appendix \ref{sec:decoherence}.
+
+\subsection{Transverse-field Hamiltonian}
+
+The transverse-field (TVF) twisting Hamiltonian with unit twisting
+strength is
+\begin{align}
+  H_{\t{TVF}}
+  = S_z^2 - \Omega S_x
+  = S_z^2 - \f{\Omega}{2} S_+ - \f{\Omega}{2} S_-.
+\end{align}
+Unlike in the case of OAT, we now cannot divide all operators of the
+form $S_+^\ell S_-^m S_z^n$ into independent sectors of constant
+imbalances $\ell,m$.  We therefore need to keep track of all operators
+in one vector $\v V_{\t{TVF}}$, which in turn requires imposing an
+ordering on all operators of the form $S_+^\ell S_-^m S_z^n$.  Such an
+ordering is provided by the one-to-one map
+$p:\mathbb{N}_0^3\to\mathbb{N}_0$, where $\mathbb{N}_0$ denotes the
+set of non-negative integers, given by
+\begin{align}
+  v\p{\ell,m,n} = B\p{T_{\ell mn}} + R\p{T_{\ell mn},\ell} + m,
+  \label{eq:vec_map}
+\end{align}
+where
+\begin{align}
+  T_{\ell mn} \equiv \ell + m + n,
+  &&
+  B\p{T} \equiv \f16 T \p{T+1} \p{T+2},
+  &&
+  R\p{T,\ell} \equiv \f12 \ell \p{2 T - \ell + 3}.
+\end{align}
+This map is designed to sort integers $\p{\ell,m,n}$ first by their
+sum $T_{\ell mn}$ (i.e.~the total weight of $S_+^\ell S_-^m S_z^n$),
+and then in turn by the values of $\ell$, $m$, and $n$, such that
+$v\p{\ell,m,n}<v\p{\ell',m',n'}$ if
+\begin{enumerate*}[label=(\roman*)]
+\item $T_{\ell mn}<T_{\ell'm'n'}$,
+\item $T_{\ell mn}=T_{\ell'm'n'}$ and $\ell<\ell'$,
+\item $\p{T_{\ell mn},\ell}=\p{T_{\ell'm'n'},\ell'}$ and $m<m'$, or
+\item $\p{T_{\ell mn},\ell,m}=\p{T_{\ell'm'n'},\ell',m'}$ and $n<n'$.
+\end{enumerate*}
+Fixing a maximal operator weight $M$, the vector keeping track of all
+operators is then
+\begin{align}
+  \v V_{\t{TVF}}
+  \equiv \sum_{\ell+m+n\le M} S_+^\ell S_-^m S_z^n \uv x_{v\p{\ell,m,n}}.
+\end{align}
+
+\subsection{Two-axis twisting in-plane with the decoherence axis}
+
+The two-axis twisting Hamiltonian one axis oriented along the
+decoherence axis is
+\begin{align}
+  H_{\t{TAT}}^{z,y}
+  = S_z^2 - S_y^2
+  = S_z^2 + \f14 S_+^2 + \f14 S_-^2 - \f12 S_+ S_- + \f12 S_z.
+\end{align}
+While upper and lower imbalances are not conserved by this
+Hamiltonians, the parity of the net imbalance, i.e.~$\ell-m~\t{mod}~2$
+for an operator $S_+^\ell S_-^m S_z^n$, is conserved.  There are
+therefore two sectors of operators which we can evolve independently
+to save computational overhead.
+
+\subsection{Two-axis twisting transverse to the decoherence axis}
+
+The two-axis twisting Hamiltonian with both axes transverse to the
+decoherence axis is
+\begin{align}
+  H_{\t{TAT}}^{y,x}
+  = S_y^2 - S_x^2
+  = - \f12 S_+^2 - \f12 S_-^2.
+\end{align}
+Just as with the previous TAT Hamiltonian, the parity of the net
+imbalance of operators is conserved by $H_{\t{TAT}}^{y,x}$, implying
+that there are two independent sectors of operators to evolve.
 
@@
 These relations can be used to inductively determine commutation
-relations between higher-weight spin operators.  In particular, we can
-find that
+relations with higher-weight spin operators.  For example, we can find
+that
 \begin{align}
@@
 considerably more difficult than the cases we have worked out thus
-far.  We therefore save this case for Appendix
+far.  We therefore save this work for Appendix
 \ref{sec:comm_transverse}.
 
-\section{Decoherence of Heisenberg spin operators}
-\label{sec:decoherence}
-
-The decoherence operator for uncorrelated incoherent decay ($\mu=-1$)
-or excitation ($\mu=1$) of individual spins is, with $\bmu\equiv-\mu$,
-\begin{align}
-  \D_\mu\p{\O}
-  \equiv \sum_j\p{\sigma_j^\bmu \O \sigma_j^\mu
-    - \f12\sp{\sigma_j^\bmu \sigma_j^\mu,\O}_+},
-\end{align}
-where
-\begin{align}
-  \sum_j \f12\sp{\sigma_j^\bmu \sigma_j^\mu,\O}_+
-  = \f12\sp{S-\mu S_z,\O}_+
-  = S\O - \f{\mu}{2}\sp{S_z,\O}_+
-  = \O\p{S-\mu S_z} - \f{\mu}{2}\sp{S_z,\O}_-,
-\end{align}
-so
-\begin{align}
-  \D_\mu\p{\O}
-  = \sum_j \sigma_j^\bmu \O \sigma_j^\mu
-  - \O\p{S - \mu S_z} + \f{\mu}{2} \sp{S_z,\O}_-.
-\end{align}
-We are interested in the effect of this decoherence on the dynamics of
-collective spin operators of the form $S_\bmu^\ell S_\mu^m S_z^n$.  We
-thus expand
-\begin{align}
-  \D_\mu\p{S_\bmu^\ell S_\mu^m S_z^n}
-  = \sum_j \sigma_j^\bmu S_\bmu^\ell S_\mu^m S_z^n \sigma_j^\mu
-  - S_\bmu^\ell S_\mu^m S_z^n \p{S - \mu S_z}
-  + \f{\mu}{2} \sp{S_z, S_\bmu^\ell S_\mu^m S_z^n}_-.
-  \label{eq:D_z+-}
-\end{align}
-The commutator in \eqref{eq:D_z+-} is straightforward to simplify
-using \eqref{eq:comm_z_mu}:
-\begin{align}
-  \mu \sp{S_z, S_\bmu^\ell S_\mu^m S_z^n}_-
-  = \mu \sp{S_z, S_\bmu^\ell S_\mu^m}_- S_z^n
-  = \mu\bmu \p{\ell-m} S_\bmu^\ell S_\mu^m S_z^n
-  = - \p{\ell-m} S_\bmu^\ell S_\mu^m S_z^n,
-\end{align}
-while the first term in \eqref{eq:D_z+-} can be written in terms of
-collective spin operators if we use
-\eqref{eq:comm_z_mu_base}-\eqref{eq:comm_mu_base} to push the jump
-operators together:
-\begin{align}
-  \sum_j \sigma_j^\bmu S_\bmu^\ell S_\mu^m S_z^n \sigma_j^\mu
-  = \sum_j S_\bmu^\ell \sigma_j^\bmu \sigma_j^\mu S_\mu^m \p{\mu+S_z}^n
-  = S_\bmu^\ell \p{S - \mu S_z} S_\mu^m
-  \sum_{k=0}^n \mu^{n-k} { n \choose k } S_z^k.
-\end{align}
-Finally, we need to push the middle $S_z$ to the right, which we can
-do using \eqref{eq:push_z_mu_single}:
-\begin{align}
-  \sum_j \sigma_j^\bmu S_\bmu^\ell S_\mu^m S_z^n \sigma_j^\mu
-  = S_\bmu^\ell S_\mu^m \p{S - m - \mu S_z}
-  \sum_{k=0}^n \mu^{n-k} { n \choose k } S_z^k.
-\end{align}
-Altogether, we have that
-\begin{align}
-  \D_\mu\p{S_\bmu^\ell S_\mu^m S_z^n}
-  &= S_\bmu^\ell S_\mu^m \sp{\p{S - m - \mu S_z} \sum_{k=0}^n
-    \mu^{n-k} { n \choose k } S_z^k
-    - \p{S - \mu S_z + \f12\sp{\ell-m}} S_z^n} \\
-  &= S_\bmu^\ell S_\mu^m \sp{\p{S - m - \mu S_z} \sum_{k=0}^{n-1}
-    \mu^{n-k} { n \choose k } S_z^k - \f12 \p{\ell+m} S_z^n}.
-\end{align}
-
-\section{Commutators between high-weight transverse spin operators}
+\section{Commutator between powers of two transverse spin operators}
 \label{sec:comm_transverse}
 
-To determine commutation relations between high-weight operators
-transverse to the decoherence (i.e.~$z$) axis, we compute (with
-$\bmu\equiv-\mu\in\set{+1,-1}$)
+To determine commutation relations between high-weight collective spin
+operators transverse to the decoherence (i.e.~$z$) axis, we compute
+(with $\bmu\equiv-\mu\in\set{+1,-1}$)
 \begin{align}
@@
   &= \sum_{q=0}^p \p{-1}^{p-q} { p \brack q } \sum_{\ell=0}^q
-  { q \choose \ell } \p{2\mu S_z}^\ell \p{m-n}^{q-\ell} \\
-  &= \sum_{\ell=0}^p \p{2\mu}^\ell S_z^\ell
-  \sum_{q=\ell}^p \p{-1}^{p-q}
-  { p \brack q } { q \choose \ell } \p{m-n}^{q-\ell}.
+  { q \choose \ell } \p{m-n}^{q-\ell} \p{2\mu S_z}^\ell \\
+  &= \sum_{\ell=0}^p \p{2\mu}^\ell \sum_{q=\ell}^p \p{-1}^{p-q}
+  { p \brack q } { q \choose \ell } \p{m-n}^{q-\ell} S_z^\ell.
 \end{align}
@@
   = p! { m \choose p } { n \choose p }
-  S_\bmu^{n-p} S_\mu^{m-p}
-  \sum_{\ell=0}^p \p{2\mu}^\ell \epsilon_{mnp}^{(\ell)} S_z^\ell,
+  S_\bmu^{n-p} S_\mu^{m-p} \sum_{\ell=0}^p
+  \mu^\ell \epsilon_{mnp}^{(\ell)} S_z^\ell,
 \end{align}
@@
   \epsilon_{mnp}^{(\ell)}
-  \equiv \sum_{q=\ell}^p \p{-1}^{p-q}
+  \equiv 2^\ell \sum_{q=\ell}^p \p{-1}^{p-q}
   { p \brack q } { q \choose \ell } \p{m-n}^{q-\ell},
@@
   p! { m \choose p } { n \choose p } S_\bmu^{n-p} S_\mu^{m-p}
-  \sum_{\ell=0}^p \p{2\mu}^\ell \epsilon_{mnp}^{(\ell)} S_z^\ell.
+  \sum_{\ell=0}^p \mu^\ell \epsilon_{mnp}^{(\ell)} S_z^\ell,
   \label{eq:comm_mu}
+\end{align}
+and
+\begin{align}
+  S_\mu^m S_\bmu^n
+  = \sum_{p=0}^{\min\set{m,n}}
+  p! { m \choose p } { n \choose p } S_\bmu^{n-p} S_\mu^{m-p}
+  \sum_{\ell=0}^p \mu^\ell \epsilon_{mnp}^{(\ell)} S_z^\ell.
+  \label{eq:push_mu_nu}
 \end{align}
@@
 \end{align}
-then commute $S_\bmu^q,S_\mu^r$ to expand
+then reorder $S_\bmu^q S_\mu^r$ with \eqref{eq:push_mu_nu} to expand
 \begin{align}
   S_\mu^p S_\bmu^q S_\mu^r S_\bmu^s
-  = S_\mu^{p+r} S_\bmu^{q+s} + \sum_{k=1}^{\min\set{q,r}}
+  = \sum_{k=0}^{\min\set{q,r}}
   k! { q \choose k } { r \choose k } S_\mu^{p+r-k} S_\bmu^{q-k}
-  \sum_{\ell=0}^k \p{-2\mu}^\ell \epsilon_{qrk}^{(\ell)}
+  \sum_{\ell=0}^k \p{-\mu}^\ell \epsilon_{qrk}^{(\ell)}
   S_z^\ell S_\bmu^s,
@@
   S_\mu^{p+r-k} S_\bmu^{q+s-k}
-  \sum_{\ell=0}^k \p{-2\mu}^\ell \epsilon_{qrk}^{(\ell)}
+  \sum_{\ell=0}^k \p{-\mu}^\ell \epsilon_{qrk}^{(\ell)}
   Z_{abrs;\mu}^{(\ell)},
@@
   Z_{abrs;\mu}^{(\ell)}
-  \equiv \p{-\mu s+S_z}^\ell \p{\mu\sp{r-s}+S_z}^a S_z^b
-  = \sum_{m=0}^\ell { \ell \choose m } \p{-\mu s}^{\ell-m}
+  &\equiv \p{-\mu s+S_z}^\ell \p{\mu\sp{r-s}+S_z}^a S_z^b \\
+  &= \sum_{m=0}^\ell { \ell \choose m } \p{-\mu s}^{\ell-m}
   \sum_{n=0}^a { a \choose n } \p{\mu\sp{r-s}}^{a-n} S_z^{b+m+n}.
@@
 
+\section{Decoherence of collective spin operators through individual
+  spin decay}
+\label{sec:decoherence}
+
+The decoherence operator for uncorrelated decay ($\mu=-1$) or
+excitation ($\mu=1$) of individual spins is, with $\bmu\equiv-\mu$,
+\begin{align}
+  \D_\mu\p{\O}
+  \equiv \sum_j\p{\sigma_j^\bmu \O \sigma_j^\mu
+    - \f12\sp{\sigma_j^\bmu \sigma_j^\mu,\O}_+},
+\end{align}
+where
+\begin{align}
+  \sum_j \f12\sp{\sigma_j^\bmu \sigma_j^\mu,\O}_+
+  = \f12\sp{S-\mu S_z,\O}_+
+  = S\O - \f{\mu}{2}\sp{S_z,\O}_+
+  = \O\p{S-\mu S_z} - \f{\mu}{2}\sp{S_z,\O}_-,
+\end{align}
+so
+\begin{align}
+  \D_\mu\p{\O}
+  = \sum_j \sigma_j^\bmu \O \sigma_j^\mu
+  - \O\p{S - \mu S_z} + \f{\mu}{2} \sp{S_z,\O}_-.
+\end{align}
+
+\subsection{The simple case}
+
+If we consider only ordered collective spin operators, there are two
+cases to work out: the effect of decoherence through $\D_\mu$ on
+$S_\mu^\ell S_\bmu^m S_z^n$, and the effect on
+$S_\bmu^\ell S_\mu^m S_z^n$.  We work out the latter case first,
+namely
+\begin{align}
+  \D_\mu\p{S_\bmu^\ell S_\mu^m S_z^n}
+  = \sum_j \sigma_j^\bmu S_\bmu^\ell S_\mu^m S_z^n \sigma_j^\mu
+  - S_\bmu^\ell S_\mu^m S_z^n \p{S - \mu S_z}
+  + \f{\mu}{2} \sp{S_z, S_\bmu^\ell S_\mu^m S_z^n}_-.
+  \label{eq:D_mu_nu}
+\end{align}
+The commutator in \eqref{eq:D_mu_nu} is straightforward to simplify
+using \eqref{eq:comm_z_mu}:
+\begin{align}
+  \mu \sp{S_z, S_\bmu^\ell S_\mu^m S_z^n}_-
+  = \mu \sp{S_z, S_\bmu^\ell S_\mu^m}_- S_z^n
+  = \mu\bmu \p{\ell-m} S_\bmu^\ell S_\mu^m S_z^n
+  = - \p{\ell-m} S_\bmu^\ell S_\mu^m S_z^n,
+\end{align}
+while the first term in \eqref{eq:D_mu_nu} can be written in terms of
+collective spin operators if we use \eqref{eq:comm_z_mu_base} to push
+the jump operators together, noting that $\sigma_j^\mu$ commutes with
+$S_\mu$:
+\begin{align}
+  \sum_j \sigma_j^\bmu S_\bmu^\ell S_\mu^m S_z^n \sigma_j^\mu
+  = \sum_j S_\bmu^\ell \sigma_j^\bmu \sigma_j^\mu S_\mu^m \p{\mu+S_z}^n
+  = S_\bmu^\ell \p{S - \mu S_z} S_\mu^m
+  \sum_{k=0}^n \mu^{n-k} { n \choose k } S_z^k.
+  \label{eq:combine_jumps}
+\end{align}
+Finally, we push the middle $S_z$ to the right using
+\eqref{eq:push_z_mu_single}:
+\begin{align}
+  \sum_j \sigma_j^\bmu S_\bmu^\ell S_\mu^m S_z^n \sigma_j^\mu
+  = S_\bmu^\ell S_\mu^m \p{S - m - \mu S_z}
+  \sum_{k=0}^n \mu^{n-k} { n \choose k } S_z^k.
+\end{align}
+Altogether, we have that
+\begin{align}
+  \D_\mu\p{S_\bmu^\ell S_\mu^m S_z^n}
+  &= S_\bmu^\ell S_\mu^m \sp{\p{S - m - \mu S_z} \sum_{k=0}^n
+    \mu^{n-k} { n \choose k } S_z^k
+    - \p{S - \mu S_z + \f12\sp{\ell-m}} S_z^n} \\
+  &= S_\bmu^\ell S_\mu^m \sp{\p{S - m - \mu S_z} \sum_{k=0}^{n-1}
+    \mu^{n-k} { n \choose k } S_z^k - \f12 \p{\ell+m} S_z^n}.
+\end{align}
+Decoherence through $\D_\mu$ only couples operators of the form
+$S_\bmu^\ell S_\mu^m S_z^n$ to operators of the (almost identical)
+form $S_\bmu^\ell S_\mu^m S_z^{n'}$ with $n'\le n$.
+
+\subsection{The harder case}
+
+The second case to consider for the decoherence of general ordered
+collective spin operators is
+\begin{align}
+  \D_\mu\p{S_\mu^\ell S_\bmu^m S_z^n}
+  = \sum_j \sigma_j^\bmu S_\mu^\ell S_\bmu^m S_z^n \sigma_j^\mu
+  - S_\mu^\ell S_\bmu^m S_z^n \p{S - \mu S_z}
+  + \f{\mu}{2} \sp{S_z, S_\mu^\ell S_\bmu^m S_z^n}_-.
+  \label{eq:D_mu_mu}
+\end{align}
+Similarly to before, the commutator in \eqref{eq:D_mu_mu} is
+\begin{align}
+  \mu \sp{S_z, S_\mu^\ell S_\bmu^m S_z^n}_-
+  = \mu \sp{S_z, S_\mu^\ell S_\bmu^m}_- S_z^n
+  = \mu\mu \p{\ell-m} S_\mu^\ell S_\bmu^m S_z^n
+  = \p{\ell-m} S_\mu^\ell S_\bmu^m S_z^n.
+\end{align}
+The difficulty now is with the first term in \eqref{eq:D_mu_mu}, as
+$\sigma_j^\bmu$ and $S_\mu$ do not commute.
+
 \end{document}

### 2018-08-25 soc_squeezing/cumulant_expansions.tex
added explenation of operator sorting in TAT-zy case

@@
 
+\newcommand{\N}{\mathbb{N}}
 \newcommand{\Z}{\mathbb{Z}}
@@
 \newcommand{\bmu}{{\bar\mu}}
+\newcommand{\floor}[1]{\left\lfloor{#1}\right\rfloor}
 
@@
   &&
-  H_{\t{TAT}}^{z,y} = \chi \p{S_z^2 - S_y^2},
+  H_{\t{TAT}}^{\p{z,y}} = \chi \p{S_z^2 - S_y^2},
   &&
-  H_{\t{TVF}} = \chi S_z^2 - \Omega S_x,
+  H_{\t{TVF}} = \chi S_z^2 + \Omega S_x,
 \end{align}
@@
 al.~in ref.~\cite{foss-feig2013nonequilibrium}.  Adapting exact
-expectation values with respect to the initial state $\ket{X}$ which
-has all spins pointing in $x$, in terms of $\mu,\nu\in\set{+1,-1}$ we
-find that
+expectation values with respect to the initial state $\ket{X}$
+satisfying $S_x\ket{X}=S\ket{X}$, in terms of $\mu,\nu\in\set{+1,-1}$
+we find that
 \begin{align}
@@
 and
-\begin{align}
+\begin{multline}
   \f{d}{dt} \p{S_\mu S_\nu S_z^n}
-  &= i\chi S_\mu S_\nu \p{\sp{\mu+\nu}^2 + 2 \sp{\mu+\nu} S_z} S_z^n
-  \tag*{} \\
-  &\quad + \gamma S_\mu S_\nu
+  = i\chi S_\mu S_\nu \p{\sp{\mu+\nu}^2 + 2 \sp{\mu+\nu} S_z} S_z^n \\
+  + \gamma S_\mu S_\nu
   \sp{\p{S - \delta_{\mu,-1} - \delta_{\nu,-1} + S_z} \sum_{k=0}^{n-1}
     \p{-1}^{n-k} { n \choose k } S_z^k - S_z^n},
-  \eqref{eq:evolution_OAT_pair}
-\end{align}
+  \label{eq:evolution_OAT_pair}
+\end{multline}
 for $\mu\ge\nu$.  The initial conditions for these systems of
@@
 \begin{align}
-  S_\mu = \sum_{m=-S}^S \sqrt{\p{S-\mu m}\p{S+\mu m+1}} \op{m+\mu}{m},
+  S_z^n = \sum_{m=-S}^S m^n \op{m},
   &&
-  S_z^n = \sum_{m=-S}^S m^n \op{m},
+  S_\mu = \sum_{m=-S}^S g_\mu\p{m} \op{m+\mu}{m},
+\end{align}
+with
+\begin{align}
+  g_\mu\p{S,m} \equiv \sqrt{\p{S-\mu m}\p{S+\mu m+1}},
+  \label{eq:transverse_elem}
 \end{align}
@@
   \bk{S_\mu S_z^n}_{t=0}
-  &= 2^{-N} \sum_{m=-S}^S m^n
-  \sp{{ N \choose S+m } { N \choose S+m+\mu }
-    \p{S-\mu m}\p{S+\mu m+1}}^{1/2} \\
-  &= 2^{-N} \sum_{m=-S}^S m^n { N \choose S+m } \abs{m - \mu S},
+  = 2^{-N} \sum_{m=-S}^S m^n { N \choose S+m } \abs{m - \mu S},
 \end{align}
 and
-\begin{multline}
+\begin{align}
   \bk{S_\mu S_\nu S_z^n}_{t=0}
-  = 2^{-N} \sum_{m=-S}^S m^n
-  \sp{{ N \choose S+m } { N \choose S+m+\mu+\nu }}^{1/2} \\
-  \times \sp{\p{S-\nu\sp{m+\nu}}\p{S+\mu\sp{m+\nu}+1}
-    \p{S-\nu m}\p{S+\nu m+1}}^{1/2}.
-\end{multline}
+  = 2^{-N} \sum_{m=-S+\delta_{\mu,-1}+\delta_{\nu,-1}}
+  ^{S-\delta_{\mu,1}-\delta_{\nu,1}}
+  m^n \sp{{ N \choose S+m+\mu+\nu } { N \choose S+m }}^{1/2}
+  g_\mu\p{S,m+\nu} g_\nu\p{S,m}.
+\end{align}
 
@@
 where the commutator is given in Appendix \ref{sec:comm_general}, and
-the decoherence operator in Appendix \ref{sec:decoherence}.
+the decoherence operator in Appendix \ref{sec:decoherence}.  We will
+be considering evolution of the initial states
+$\ket{X}:S_x\ket{X}=S\ket{X}$ and $\ket{-Z}:S_z\ket{-Z}=-S\ket{-Z}$,
+for which initial conditions are given by
+\begin{multline}
+  \bk{X|S_+^\ell S_-^m S_z^n|X}
+  =  2^{-N} \sum_{k=-S+m}^{S-\ell} k^n
+  \sp{{ N \choose S+k+\ell-m } { N \choose S+k }}^{1/2} \\
+  \times \prod_{p=0}^{\ell-1} g_+\p{S,k-m+p}
+  \prod_{q=0}^{m-1} g_-\p{S,k-q}
+\end{multline}
+and
+\begin{multline}
+  \bk{-Z|S_+^\ell S_-^m S_z^n|-Z}
+  = \p{-S}^n 2^{-N} \sum_{k=-S+m}^{S-\ell}
+  \sp{{ N \choose S+k+\ell-m } { N \choose S+k }}^{1/2} \\
+  \times \prod_{p=0}^{\ell-1} g_+\p{S,k-m+p}
+  \prod_{q=0}^{m-1} g_-\p{S,k-q}.
+\end{multline}
 
@@
   H_{\t{TVF}}
-  = S_z^2 - \Omega S_x
-  = S_z^2 - \f{\Omega}{2} S_+ - \f{\Omega}{2} S_-.
+  = S_z^2 + \Omega S_x
+  = S_z^2 + \f{\Omega}{2} S_+ + \f{\Omega}{2} S_-.
 \end{align}
@@
 ordering on all operators of the form $S_+^\ell S_-^m S_z^n$.  Such an
-ordering is provided by the one-to-one map
-$p:\mathbb{N}_0^3\to\mathbb{N}_0$, where $\mathbb{N}_0$ denotes the
-set of non-negative integers, given by
+ordering is provided by the one-to-one map $u:\N_0^3\to\N_0$, where$
+\N_0$ denotes the set of non-negative integers, given by
 \begin{align}
-  v\p{\ell,m,n} = B\p{T_{\ell mn}} + R\p{T_{\ell mn},\ell} + m,
+  u\p{\ell,m,n} = B\p{T_{\ell mn}} + R\p{T_{\ell mn},\ell} + m,
   \label{eq:vec_map}
 \end{align}
-where
+where $T_{\ell mn}\equiv\ell+m+n$ for shorthand and
 \begin{align}
-  T_{\ell mn} \equiv \ell + m + n,
-  &&
-  B\p{T} \equiv \f16 T \p{T+1} \p{T+2},
-  &&
-  R\p{T,\ell} \equiv \f12 \ell \p{2 T - \ell + 3}.
+  B\p{T}
+  \equiv \abs{\set{\p{\ell,m,n}\in\N_0^3 : T_{\ell mn}<T}}
+  = \f16 T \p{T+1} \p{T+2},
 \end{align}
-This map is designed to sort integers $\p{\ell,m,n}$ first by their
-sum $T_{\ell mn}$ (i.e.~the total weight of $S_+^\ell S_-^m S_z^n$),
-and then in turn by the values of $\ell$, $m$, and $n$, such that
-$v\p{\ell,m,n}<v\p{\ell',m',n'}$ if
-\begin{enumerate*}[label=(\roman*)]
-\item $T_{\ell mn}<T_{\ell'm'n'}$,
-\item $T_{\ell mn}=T_{\ell'm'n'}$ and $\ell<\ell'$,
-\item $\p{T_{\ell mn},\ell}=\p{T_{\ell'm'n'},\ell'}$ and $m<m'$, or
-\item $\p{T_{\ell mn},\ell,m}=\p{T_{\ell'm'n'},\ell',m'}$ and $n<n'$.
-\end{enumerate*}
-Fixing a maximal operator weight $M$, the vector keeping track of all
-operators is then
+\begin{align}
+  R\p{T,L}
+  \equiv \abs{\set{\p{\ell,m,n}\in\N_0^3 : T_{\ell mn}=T , \ell<L}}
+  = \f12 \ell \p{2 T - \ell + 3}.
+\end{align}
+In words, $u\p{\ell,m,n}$ sorts the collection non-negative integers
+$\p{\ell,m,n}$ first by their sum $T_{\ell mn}$, and then in turn by
+the values of $\ell$ and $m$.  Fixing a maximal operator weight $M$,
+the vector keeping track of all operators is then
 \begin{align}
   \v V_{\t{TVF}}
-  \equiv \sum_{\ell+m+n\le M} S_+^\ell S_-^m S_z^n \uv x_{v\p{\ell,m,n}}.
+  \equiv \sum_{\ell+m+n\le M} S_+^\ell S_-^m S_z^n \uv x_{u\p{\ell,m,n}}.
 \end{align}
@@
 \begin{align}
-  H_{\t{TAT}}^{z,y}
+  H_{\t{TAT}}^{\p{z,y}}
   = S_z^2 - S_y^2
@@
 therefore two sectors of operators which we can evolve independently
-to save computational overhead.
+to save computational overhead.  To keep track of these operators in
+one-dimensional vectors, we define the onto maps $v_p:\N_0^3\to\N_0$
+for $p\in\set{0,1}$ given by
+\begin{align}
+  v_p\p{\ell,m,n} = B_p\p{T_{\ell mn}} + R_p\p{T_{\ell mn},n} + m,
+\end{align}
+where
+\begin{align}
+  B_p\p{T}
+  &\equiv \abs{\set{\p{\ell,m,n}\in\N_0^3 :
+      T_{\ell mn}<T, \ell-m=p~\t{mod}~2}} \\
+  &= \f12 B\p{T} + \f14 \p{T+1} \p{T~\t{mod}~2} \p{1-2p},
+\end{align}
+\begin{align}
+  R_p\p{T,N}
+  &\equiv \abs{\set{\p{\ell,m,n}\in\N_0^3 :
+      T_{\ell mn}=T, \ell-m=p~\t{mod}~2, n < N }} \\
+  &= \floor{\f{n}{2}}
+  \p{T - \floor{\f{n}{2}} + 2 - \p{T+p~\t{mod}~2}}.
+\end{align}
+The map $v_p\p{\ell,m,n}$ sorts the collection non-negative integers
+$\p{\ell,m,n}$ with $\ell-m=p~\t{mod}~2$ first by their sum
+$T_{\ell mn}$, and then in turn by the values of $n$ and $m$.  Fixing
+a maximum operator weight $M$, the vectors keeping track of all
+operators are then
+\begin{align}
+  \v V_{\t{TAT}}^{\p{z,y,p}}
+  \equiv \sum_{\substack{\ell,m,n\le M\\\ell-m=p~\t{mod}~2}}
+  S_+^\ell S_-^m S_z^n \uv x_{v_p\p{\ell,m,n}}.
+\end{align}
 

### 2018-08-27 soc_squeezing/cumulant_expansions.tex
replace explanations for dynamics with TVF, TAT by that for the general case

@@
   &&
-  H_{\t{TAT}}^{\p{z,y}} = \chi \p{S_z^2 - S_y^2},
+  H_{\t{TAT}}^{z,y} = \chi \p{S_z^2 - S_y^2},
   &&
@@
     \p{-1}^{n-k} { n \choose k } S_z^k - \f12 S_z^n}
-  \label{eq:evolution_OAT_single}
 \end{align}
-and
-\begin{multline}
-  \f{d}{dt} \p{S_\mu S_\nu S_z^n}
-  = i\chi S_\mu S_\nu \p{\sp{\mu+\nu}^2 + 2 \sp{\mu+\nu} S_z} S_z^n \\
-  + \gamma S_\mu S_\nu
-  \sp{\p{S - \delta_{\mu,-1} - \delta_{\nu,-1} + S_z} \sum_{k=0}^{n-1}
+for $\mu\in\set{+1,-1}$, and
+\begin{align}
+  \f{d}{dt} \p{S_+^2 S_z^n}
+  = 4i\chi S_\mu S_\nu \p{1 + S_z} S_z^n
+  + \gamma S_\mu S_\nu \sp{\p{S + S_z} \sum_{k=0}^{n-1}
     \p{-1}^{n-k} { n \choose k } S_z^k - S_z^n},
-  \label{eq:evolution_OAT_pair}
-\end{multline}
-for $\mu\ge\nu$.  The initial conditions for these systems of
-equations can be determined from the expansions
+\end{align}
+while $d/dt\p{S_\pm S_\mp}=0$, so
+\begin{align}
+  \bk{S_\pm S_\mp} = \bk{S_\pm S_\mp}_{t=0} e^{-\gamma t}.
+\end{align}
+The initial conditions for these systems of equations can be
+determined from the expansions
 \begin{align}
@@
 
-\subsection{Operator imbalance, conservation laws, and time evolution}
-
-For convenience of language, for collective spin operators of the form
-$S_+^\ell S_-^m S_z^n$ we respectively define the {\it upper}, {\it
-  lower}, and {\it net imbalance}
-\begin{align}
-  \I_+\p{S_+^\ell S_-^m S_z^n} \equiv \ell,
-  &&
-  \I_-\p{S_+^\ell S_-^m S_z^n} \equiv m,
-  &&
-  \I\p{S_+^\ell S_-^m S_z^n} \equiv \ell - m.
-\end{align}
-In the case of one-axis twisting, both upper and lower (and net)
-imbalances are {\it conserved}, meaning that operators with upper and
-lower imbalances $\p{\ell,m}$ depend only on other operators with the
-same imbalances.  We can therefore divide the task of operator time
-evolution into sectors of constant imbalances, which greatly reduces
-computational overhead.  Specifically, operator time evolution now
-involves evolving ``operator vectors''
-\begin{align}
-  \v V_{\t{OAT}}^{(\ell,m)}
-  \equiv S_+^\ell S_-^m \sum_{n\ge0} S_z^n \uv x_n,
-\end{align}
-which in practice must be truncated at some maximal $n=M$.  In the
-case of OAT, the time derivative of $S_+^\ell S_-^m S_z^n$ depends on
-$S_+^\ell S_-^m S_z^{n+1}$ which implies that we can only trust the
-value of $S_+^\ell S_-^m S_z^n$ up to the time at which the initial
-value of $S_+^\ell S_-^m S_z^M$ has (conservatively) an $\O\p{1}$
-contribution to $S_+^\ell S_-^m S_z^n$.
-
-In order to compute squeezing from OAT, we need only need to evolve
-$\v V_{\t{OAT}}^{(1,0)}$, $\v V_{\t{OAT}}^{(0,1)}$,
-$\v V_{\t{OAT}}^{(1,1)}$, and $\v V_{\t{OAT}}^{(2,0)}$.  The values of
-axial spin operators $S_z$ and $S_z^2$ are known exactly at all times
-from \eqref{eq:Sz_OAT}.
-
-\section{Transverse-field and two-axis twisting}
+\section{The general case}
 
@@
 \end{align}
+The Hamiltonian for one-axis twisting with a transverse-field, for
+example, is
+\begin{align}
+  H_{\t{TVF}}
+  = \chi S_z^2 + \Omega S_x
+  = \chi S_z^2 + \f{\Omega}{2} S_+ + \f{\Omega}{2} S_-,
+\end{align}
+while the two-axis twisting Hamiltonians about $\p{z,y}$ and $\p{y,x}$
+are
+\begin{align}
+  H_{\t{TAT}}^{z,y}
+  = \chi \p{S_z^2 - S_y^2}
+  = \chi \p{S_z^2 + \f14 S_+^2 + \f14 S_-^2 - \f12 S_+ S_- + \f12 S_z},
+\end{align}
+and
+\begin{align}
+  H_{\t{TAT}}^{y,x}
+  = \chi\p{S_y^2 - S_x^2}
+  = -\f{\chi}{2} \p{S_+^2 + S_-^2}.
+\end{align}
 The evolution of a general operator $S_+^\ell S_-^m S_z^n$ under a
@@
   = i\sum_{p,q,r} h_{pqr} \sp{S_+^p S_-^q S_z^r, S_+^\ell S_-^m S_z^n}_-
-  + \gamma \D_-\p{S_+^\ell S_-^m S_z^n},
+  + \gamma \D_-\p{S_+^\ell S_-^m S_z^n}
+  = \sum_{p,q,r} T^{\ell m n}_{pqr} S_+^p S_-^q S_z^r,
 \end{align}
@@
 \end{multline}
-
-\subsection{Transverse-field Hamiltonian}
-
-The transverse-field (TVF) twisting Hamiltonian with unit twisting
-strength is
+Equipped with these initial conditions, we can expand
 \begin{align}
-  H_{\t{TVF}}
-  = S_z^2 + \Omega S_x
-  = S_z^2 + \f{\Omega}{2} S_+ + \f{\Omega}{2} S_-.
+  \bk{S_+^\ell S_-^m S_z^n}
+  = \sum_{k\ge0} \f{t^k}{k!}
+  \bk{\f{d^k}{dt^k}\p{S_+^\ell S_-^m S_z^n}}_{t=0}
+  \approx \sum_{k=0}^M \f{t^k}{k!}
+  \bk{\f{d^k}{dt^k}\p{S_+^\ell S_-^m S_z^n}}_{t=0},
 \end{align}
-Unlike in the case of OAT, we now cannot divide all operators of the
-form $S_+^\ell S_-^m S_z^n$ into independent sectors of constant
-imbalances $\ell,m$.  We therefore need to keep track of all operators
-in one vector $\v V_{\t{TVF}}$, which in turn requires imposing an
-ordering on all operators of the form $S_+^\ell S_-^m S_z^n$.  Such an
-ordering is provided by the one-to-one map $u:\N_0^3\to\N_0$, where$
-\N_0$ denotes the set of non-negative integers, given by
+where $M$ is chosen such that
 \begin{align}
-  u\p{\ell,m,n} = B\p{T_{\ell mn}} + R\p{T_{\ell mn},\ell} + m,
-  \label{eq:vec_map}
+  \f{t^M}{M!} \bk{\f{d^M}{dt^M}\p{S_+^\ell S_-^m S_z^n}}_{t=0}
+  \ll S^{\ell+m+n}.
 \end{align}
-where $T_{\ell mn}\equiv\ell+m+n$ for shorthand and
-\begin{align}
-  B\p{T}
-  \equiv \abs{\set{\p{\ell,m,n}\in\N_0^3 : T_{\ell mn}<T}}
-  = \f16 T \p{T+1} \p{T+2},
-\end{align}
-\begin{align}
-  R\p{T,L}
-  \equiv \abs{\set{\p{\ell,m,n}\in\N_0^3 : T_{\ell mn}=T , \ell<L}}
-  = \f12 \ell \p{2 T - \ell + 3}.
-\end{align}
-In words, $u\p{\ell,m,n}$ sorts the collection non-negative integers
-$\p{\ell,m,n}$ first by their sum $T_{\ell mn}$, and then in turn by
-the values of $\ell$ and $m$.  Fixing a maximal operator weight $M$,
-the vector keeping track of all operators is then
-\begin{align}
-  \v V_{\t{TVF}}
-  \equiv \sum_{\ell+m+n\le M} S_+^\ell S_-^m S_z^n \uv x_{u\p{\ell,m,n}}.
-\end{align}
-
-\subsection{Two-axis twisting in-plane with the decoherence axis}
-
-The two-axis twisting Hamiltonian one axis oriented along the
-decoherence axis is
-\begin{align}
-  H_{\t{TAT}}^{\p{z,y}}
-  = S_z^2 - S_y^2
-  = S_z^2 + \f14 S_+^2 + \f14 S_-^2 - \f12 S_+ S_- + \f12 S_z.
-\end{align}
-While upper and lower imbalances are not conserved by this
-Hamiltonians, the parity of the net imbalance, i.e.~$\ell-m~\t{mod}~2$
-for an operator $S_+^\ell S_-^m S_z^n$, is conserved.  There are
-therefore two sectors of operators which we can evolve independently
-to save computational overhead.  To keep track of these operators in
-one-dimensional vectors, we define the onto maps $v_p:\N_0^3\to\N_0$
-for $p\in\set{0,1}$ given by
-\begin{align}
-  v_p\p{\ell,m,n} = B_p\p{T_{\ell mn}} + R_p\p{T_{\ell mn},n} + m,
-\end{align}
-where
-\begin{align}
-  B_p\p{T}
-  &\equiv \abs{\set{\p{\ell,m,n}\in\N_0^3 :
-      T_{\ell mn}<T, \ell-m=p~\t{mod}~2}} \\
-  &= \f12 B\p{T} + \f14 \p{T+1} \p{T~\t{mod}~2} \p{1-2p},
-\end{align}
-\begin{align}
-  R_p\p{T,N}
-  &\equiv \abs{\set{\p{\ell,m,n}\in\N_0^3 :
-      T_{\ell mn}=T, \ell-m=p~\t{mod}~2, n < N }} \\
-  &= \floor{\f{n}{2}}
-  \p{T - \floor{\f{n}{2}} + 2 - \p{T+p~\t{mod}~2}}.
-\end{align}
-The map $v_p\p{\ell,m,n}$ sorts the collection non-negative integers
-$\p{\ell,m,n}$ with $\ell-m=p~\t{mod}~2$ first by their sum
-$T_{\ell mn}$, and then in turn by the values of $n$ and $m$.  Fixing
-a maximum operator weight $M$, the vectors keeping track of all
-operators are then
-\begin{align}
-  \v V_{\t{TAT}}^{\p{z,y,p}}
-  \equiv \sum_{\substack{\ell,m,n\le M\\\ell-m=p~\t{mod}~2}}
-  S_+^\ell S_-^m S_z^n \uv x_{v_p\p{\ell,m,n}}.
-\end{align}
-
-\subsection{Two-axis twisting transverse to the decoherence axis}
-
-The two-axis twisting Hamiltonian with both axes transverse to the
-decoherence axis is
-\begin{align}
-  H_{\t{TAT}}^{y,x}
-  = S_y^2 - S_x^2
-  = - \f12 S_+^2 - \f12 S_-^2.
-\end{align}
-Just as with the previous TAT Hamiltonian, the parity of the net
-imbalance of operators is conserved by $H_{\t{TAT}}^{y,x}$, implying
-that there are two independent sectors of operators to evolve.
+% comment on explosion of operators
 

### 2018-08-27 soc_squeezing/cumulant_expansions.tex
moved initial conditions calculations to an appendix

@@
     \p{-1}^{n-k} { n \choose k } S_z^k - \f12 S_z^n}
+  \label{eq:diff_mu_OAT}
 \end{align}
@@
     \p{-1}^{n-k} { n \choose k } S_z^k - S_z^n},
+  \label{eq:diff_pp_OAT}
 \end{align}
@@
 \end{align}
-The initial conditions for these systems of equations can be
-determined from the expansions
-\begin{align}
-  S_z^n = \sum_{m=-S}^S m^n \op{m},
-  &&
-  S_\mu = \sum_{m=-S}^S g_\mu\p{m} \op{m+\mu}{m},
-\end{align}
-with
-\begin{align}
-  g_\mu\p{S,m} \equiv \sqrt{\p{S-\mu m}\p{S+\mu m+1}},
-  \label{eq:transverse_elem}
-\end{align}
-and the initial state
-\begin{align}
-  \ket{X} = 2^{-N/2} \sum_{m=-S}^S { N \choose S+m }^{1/2} \ket{m},
-\end{align}
-from which we can determine
-\begin{align}
-  \bk{S_\mu S_z^n}_{t=0}
-  = 2^{-N} \sum_{m=-S}^S m^n { N \choose S+m } \abs{m - \mu S},
-\end{align}
-and
-\begin{align}
-  \bk{S_\mu S_\nu S_z^n}_{t=0}
-  = 2^{-N} \sum_{m=-S+\delta_{\mu,-1}+\delta_{\nu,-1}}
-  ^{S-\delta_{\mu,1}-\delta_{\nu,1}}
-  m^n \sp{{ N \choose S+m+\mu+\nu } { N \choose S+m }}^{1/2}
-  g_\mu\p{S,m+\nu} g_\nu\p{S,m}.
-\end{align}
+The initial conditions for \eqref{eq:diff_mu_OAT} and
+\eqref{eq:diff_pp_OAT} are provided in Appendix
+\ref{sec:initial_conditions}.
 
@@
 where the commutator is given in Appendix \ref{sec:comm_general}, and
-the decoherence operator in Appendix \ref{sec:decoherence}.  We will
-be considering evolution of the initial states
-$\ket{X}:S_x\ket{X}=S\ket{X}$ and $\ket{-Z}:S_z\ket{-Z}=-S\ket{-Z}$,
-for which initial conditions are given by
-\begin{multline}
-  \bk{X|S_+^\ell S_-^m S_z^n|X}
-  =  2^{-N} \sum_{k=-S+m}^{S-\ell} k^n
-  \sp{{ N \choose S+k+\ell-m } { N \choose S+k }}^{1/2} \\
-  \times \prod_{p=0}^{\ell-1} g_+\p{S,k-m+p}
-  \prod_{q=0}^{m-1} g_-\p{S,k-q}
-\end{multline}
-and
-\begin{multline}
-  \bk{-Z|S_+^\ell S_-^m S_z^n|-Z}
-  = \p{-S}^n 2^{-N} \sum_{k=-S+m}^{S-\ell}
-  \sp{{ N \choose S+k+\ell-m } { N \choose S+k }}^{1/2} \\
-  \times \prod_{p=0}^{\ell-1} g_+\p{S,k-m+p}
-  \prod_{q=0}^{m-1} g_-\p{S,k-q}.
-\end{multline}
-Equipped with these initial conditions, we can expand
+the decoherence operator in Appendix \ref{sec:decoherence}.  We can
+then expand
 \begin{align}
@@
 \end{align}
+Initial conditions, i.e.~expectation values at $t=0$, are provided in
+Appendix \ref{sec:initial_conditions}.
 % comment on explosion of operators
@@
 
+\section{Initial conditions}
+\label{sec:initial_conditions}
+
+Initial conditions for all collective spin operators can be determined
+from the expansions
+\begin{align}
+  S_z^n = \sum_{m=-S}^S m^n \op{m},
+  &&
+  S_\mu = \sum_{m=-S}^S g_\mu\p{m} \op{m+\mu}{m},
+\end{align}
+with
+\begin{align}
+  g_\mu\p{S,m} \equiv \sqrt{\p{S-\mu m}\p{S+\mu m+1}},
+\end{align}
+and the initial state
+\begin{align}
+  \ket{X} \equiv 2^{-N/2} \sum_{m=-S}^S { N \choose S+m }^{1/2} \ket{m},
+\end{align}
+using which we can compute
+\begin{multline}
+  \bk{X|S_+^\ell S_-^m S_z^n|X}
+  =  2^{-N} \sum_{k=-S}^{S-\max\set{\ell,m}} \p{k+m}^n
+  \sp{{ N \choose S+k+\ell } { N \choose S+k+m }}^{1/2} \\
+  \times \prod_{p=0}^{\ell-1} g_+\p{S,k+p}
+  \prod_{q=0}^{m-1} g_+\p{S,k-q}.
+\end{multline}
+
 \end{document}

### 2018-08-28 soc_squeezing/cumulant_expansions.tex
simplified expectation values
removed obsolete section and changed spin operator ordering

@@
 \begin{align}
-  \Phi\p{\zeta,t} \equiv e^{-\gamma t/2}
-  \sp{\cos\p{s_\zeta t}+\f12\gamma t~\t{sinc}\p{s_\zeta t}},
+  \Phi\p{\xi,t} \equiv e^{-\gamma t/2}
+  \sp{\cos\p{s_\xi t}+\f12\gamma t~\t{sinc}\p{s_\xi t}},
   &&
-  \Psi\p{\zeta,t} \equiv e^{-\gamma t/2}
-  \p{is_\zeta-\gamma/2}t~ \t{sinc}\p{s_\zeta t},
+  \Psi\p{\xi,t} \equiv e^{-\gamma t/2}
+  \p{is_\xi-\gamma/2}t~ \t{sinc}\p{s_\xi t},
 \end{align}
-and $s_\zeta \equiv \zeta + i2\gamma$.  The last ingredients we need
+and $s_\xi \equiv \xi + i2\gamma$.  The last ingredients we need
 to compute spin squeezing at any time are $\bk{S_z}$ and $\bk{S_z^2}$.
@@
 $\xi^2$ in \eqref{eq:squeezing} at any time throughout evolution of
-the initial state $\ket{X}$ under $H_{\t{OAT}}$.
+the initial state $\ket{X}:S_x\ket{X}=S\ket{X}$ under $H_{\t{OAT}}$.
 
-\subsection{Heisenberg systems of equations}
+\section{Cumulant expansion and Heisenberg systems of equations}
 
@@
 
-The equations of motion for $S_\mu$, $S_\mu S_z$, and $S_\mu S_\nu$
-(i.e.~where $\mu,\nu\in\set{+1,-1}$) under the OAT Hamiltonian
-$H_{\t{OAT}}=\chi S_z^2$ are determined by
-\begin{align}
-  \f{d}{dt} \p{S_+^\ell S_-^m S_z^n}
-  = i\chi \sp{S_z^2, S_+^\ell S_-^m S_z^n}_-
-  + \gamma \D_-\p{S_+^\ell S_-^m S_z^n},
-\end{align}
-where we can use identity \eqref{eq:comm_z_mu} to simplify
-\begin{align}
-  \sp{S_z^2, S_+^\ell S_-^m S_z^n}_-
-  = \sp{S_z^2, S_+^\ell S_-^m}_- S_z^n
-  = S_+^\ell S_-^m \p{\sp{\ell-m}^2 + 2\sp{\ell-m} S_z} S_z^n,
-  \label{eq:S_z^2_comm}
-\end{align}
-and the effect of decoherence through $\D_-\p{S_+^\ell S_-^m S_z^n}$ is
-worked out in Appendix \ref{sec:decoherence}.  In order to compute
-squeezing of the OAT Hamiltonian in the presence of single-spin decay,
-we therefore need to solve the systems of equations defined by
-\begin{align}
-  \f{d}{dt} \p{S_\mu S_z^n}
-  = i\chi S_\mu \p{1 + 2\mu S_z} S_z^n
-  + \gamma S_\mu \sp{\p{S - \delta_{\mu,-1} + S_z} \sum_{k=0}^{n-1}
-    \p{-1}^{n-k} { n \choose k } S_z^k - \f12 S_z^n}
-  \label{eq:diff_mu_OAT}
-\end{align}
-for $\mu\in\set{+1,-1}$, and
-\begin{align}
-  \f{d}{dt} \p{S_+^2 S_z^n}
-  = 4i\chi S_\mu S_\nu \p{1 + S_z} S_z^n
-  + \gamma S_\mu S_\nu \sp{\p{S + S_z} \sum_{k=0}^{n-1}
-    \p{-1}^{n-k} { n \choose k } S_z^k - S_z^n},
-  \label{eq:diff_pp_OAT}
-\end{align}
-while $d/dt\p{S_\pm S_\mp}=0$, so
-\begin{align}
-  \bk{S_\pm S_\mp} = \bk{S_\pm S_\mp}_{t=0} e^{-\gamma t}.
-\end{align}
-The initial conditions for \eqref{eq:diff_mu_OAT} and
-\eqref{eq:diff_pp_OAT} are provided in Appendix
-\ref{sec:initial_conditions}.
-
-\section{The general case}
-
 All collective spin Hamiltonians can be written in the form
 \begin{align}
-  H = \sum_{p,q,r} h_{pqr} S_+^p S_-^q S_z^r.
+  H = \sum_{p,q,r} h_{pqr} S_+^p S_z^q S_-^r.
   \label{eq:H_general}
 \end{align}
-The Hamiltonian for one-axis twisting with a transverse-field, for
+The Hamiltonians for one-axis twisting with a transverse-field, for
 example, is
@@
 \begin{align}
-  \f{d}{dt} \p{S_+^\ell S_-^m S_z^n}
-  = i\sum_{p,q,r} h_{pqr} \sp{S_+^p S_-^q S_z^r, S_+^\ell S_-^m S_z^n}_-
-  + \gamma \D_-\p{S_+^\ell S_-^m S_z^n}
-  = \sum_{p,q,r} T^{\ell m n}_{pqr} S_+^p S_-^q S_z^r,
+  \f{d}{dt} \p{S_+^\ell S_z^m S_-^n}
+  = i\sum_{p,q,r} h_{pqr} \sp{S_+^p S_z^q S_-^r, S_+^\ell S_z^m S_-^n}_-
+  + \gamma \D_-\p{S_+^\ell S_z^m S_-^n}
+  \equiv \sum_{p,q,r} T^{\ell m n}_{pqr} S_+^p S_z^q S_-^r,
 \end{align}
-where the commutator is given in Appendix \ref{sec:comm_general}, and
-the decoherence operator in Appendix \ref{sec:decoherence}.  We can
-then expand
+where the commutator is worked out in Appendix \ref{sec:comm_general},
+and the decoherence operator in Appendix \ref{sec:decoherence}.  We
+can then expand
 \begin{align}
-  \bk{S_+^\ell S_-^m S_z^n}
+  \bk{S_+^\ell S_z^m S_-^n}
   = \sum_{k\ge0} \f{t^k}{k!}
-  \bk{\f{d^k}{dt^k}\p{S_+^\ell S_-^m S_z^n}}_{t=0}
-  \approx \sum_{k=0}^M \f{t^k}{k!}
-  \bk{\f{d^k}{dt^k}\p{S_+^\ell S_-^m S_z^n}}_{t=0},
+  \bk{\f{d^k}{dt^k}\p{S_+^\ell S_z^m S_-^n}}_{t=0}
+  \approx \sum_{k=0}^{M-1} \f{t^k}{k!}
+  \bk{\f{d^k}{dt^k}\p{S_+^\ell S_z^m S_-^n}}_{t=0},
 \end{align}
@@
 \begin{align}
-  \f{t^M}{M!} \bk{\f{d^M}{dt^M}\p{S_+^\ell S_-^m S_z^n}}_{t=0}
+  \f{t^M}{M!} \bk{\f{d^M}{dt^M}\p{S_+^\ell S_z^m S_-^n}}_{t=0}
   \ll S^{\ell+m+n}.
@@
   = 2 \mu S_\mu^m + S_\mu S_z S_\mu^{m-2}
-  = S_\mu^m \p{\mu m + S_z},
+  = S_\mu^m \p{\mu m + S_z}.
   \label{eq:push_z_mu_single}
-\end{align}
-and in turn
-\begin{align}
-  S_z S_\mu^m S_\bmu^n
-  = S_\mu^m \p{\mu m + S_z} S_\bmu^n
-  = S_\mu^m S_\bmu^n \p{\mu\sp{m - n} + S_z}.
 \end{align}
@@
 \begin{align}
-  S_z^\ell S_\mu^m S_\bmu^n
-  = S_\mu^m S_\bmu^n \p{\mu\sp{m - n} + S_z}^\ell,
+  S_z^\ell S_\mu^m
+  = S_\mu^m \p{\mu m + S_z}^\ell,
   \label{eq:push_z_mu}
@@
 \begin{align}
-  \sp{S_z^\ell, S_\mu^m S_\bmu^n}_-
-  = S_\mu^m S_\bmu^n \sp{\mu^\ell \p{\mu\sp{m - n} + S_z}^\ell - S_z^\ell}
-  = S_\mu^m S_\bmu^n  \sum_{k=0}^{\ell-1} {\ell \choose k}
-  \mu^{\ell-k} \p{m-n}^{\ell-k} S_z^k.
+  \sp{S_z^\ell, S_\mu^m}_-
+  = S_\mu^m \sp{\p{\mu m + S_z}^\ell - S_z^\ell}
+  = S_\mu^m \sum_{k=0}^{\ell-1} {\ell \choose k} \p{\mu m}^{\ell-k} S_z^k.
   \label{eq:comm_z_mu}
@@
 track of high-weight operators that are linearly independent.  We
-(arbitrarily) choose to impose an ordering with all $S_z$ operators on
-the right.  This choice of ordering compels us to expand
+choose to impose an ordering for which all collective spin operators
+take the form $S_\bmu^\ell S_\mu^m S_z^n$, as such an ordering will
+prove convenient for the calculations in this section.  This choice of
+ordering compels us to expand
 \begin{align}
@@
   { p \brack q } { q \choose \ell } \p{m-n}^{q-\ell},
-  \label{eq:epsilon}
+  \label{eq:eps_mnp}
 \end{align}
@@
   p! { m \choose p } { n \choose p } S_\bmu^{n-p} S_\mu^{m-p}
-  \sum_{\ell=0}^p \mu^\ell \epsilon_{mnp}^{(\ell)} S_z^\ell.
+  \sum_{\ell=0}^p \mu^\ell \epsilon_{mnp}^{(\ell)} S_z^\ell,
   \label{eq:push_mu_nu}
 \end{align}
+If we wish to order collective spin operators with $S_z$ in between
+$S_\bmu$ and $S_\mu$, then we find
+\begin{align}
+  S_\mu^m S_\bmu^n
+  = \sum_{p=0}^{\min\set{m,n}} p! { m \choose p } { n \choose p }
+  S_\bmu^{n-p} Z_{mn;\mu}^{(p)} S_\mu^{m-p},
+\end{align}
+where
+\begin{align}
+  Z_{mn;\mu}^{(p)}
+  \equiv \sum_{\ell=0}^p \mu^\ell \epsilon_{mnp}^{(\ell)}
+  \p{-\mu\sp{m-p} + S_z}^\ell
+  = \sum_{q=0}^p \mu^q \xi_{mnp}^{(q)} S_z^q,
+  \label{eq:Z_mn}
+\end{align}
+with
+\begin{align}
+  \xi_{mnp}^{(q)}
+  \equiv \sum_{\ell=q}^p \p{-1}^{\ell-q}
+  \epsilon_{mnp}^{(\ell)} { \ell \choose q } \p{m-p}^{\ell-q}
+  \label{eq:xi_mnp}
+\end{align}
+in terms of $\epsilon_{mnp}^{(\ell)}$ as defined in
+\eqref{eq:eps_mnp}.
 
@@
 
-If all collective spin operators are written as ordered products of
-$S_+$, $S_-$, and $S_z$ with all $S_z$ on the right, then the most
-general commutator we need to consider is (with
-$\bmu\equiv-\mu\in\set{+1,-1}$)
+If all collective spin operators are written in the form
+$S_\mu^\ell S_z^m S_\bmu^n$ for $\bmu\equiv-\mu$, then the most
+general commutator we need to consider is
 \begin{align}
-  \sp{S_\mu^p S_\bmu^q S_z^a, S_\mu^r S_\bmu^s S_z^b}_-
-  = \S^{pqa}_{rsb;\mu} - \S^{rsb}_{pqa;\mu},
-  &&
-  \S^{pqa}_{rsb;\mu}
-  \equiv S_\mu^p S_\bmu^q S_z^a S_\mu^r S_\bmu^s S_z^b.
+  \sp{S_\mu^p S_z^a S_\bmu^q, S_\mu^r S_z^b S_\bmu^s}_-
+  = \S^{paq}_{rbs;\mu} - \S^{rbs}_{paq;\mu},
 \end{align}
-We can push the $S_z^a$ in $\S^{pqa}_{rsb;\mu}$ to the right through
-$S_\mu^r S_\bmu^s$ to get
+where
 \begin{align}
-  \S^{pqa}_{rsb;\mu}
-  = S_\mu^p S_\bmu^q S_\mu^r S_\bmu^s  \p{\mu\sp{r-s}+S_z}^a S_z^b,
+  \S^{paq}_{rbs;\mu}
+  \equiv S_\mu^p S_z^a S_\bmu^q S_\mu^r S_z^b S_\bmu^s
+  = \sum_{k=0}^{\min\set{q,r}} k! { q \choose k } { r \choose k }
+  S_\mu^{p+r-k} \tilde Z_{qrab;\bmu}^{(k)} S_\bmu^{q+s-k},
 \end{align}
-then reorder $S_\bmu^q S_\mu^r$ with \eqref{eq:push_mu_nu} to expand
+and, in terms of $\xi_{qrk}^{(\ell)}$ as defined in \eqref{eq:xi_mnp},
 \begin{align}
-  S_\mu^p S_\bmu^q S_\mu^r S_\bmu^s
-  = \sum_{k=0}^{\min\set{q,r}}
-  k! { q \choose k } { r \choose k } S_\mu^{p+r-k} S_\bmu^{q-k}
-  \sum_{\ell=0}^k \p{-\mu}^\ell \epsilon_{qrk}^{(\ell)}
-  S_z^\ell S_\bmu^s,
-\end{align}
-and finally push $S_z^\ell$ to the right through $S_\bmu^s$, finding
-\begin{align}
-  \S^{pqa}_{rsb;\mu}
-  = \sum_{k=0}^{\min\set{q,r}} k! { q \choose k } { r \choose k }
-  S_\mu^{p+r-k} S_\bmu^{q+s-k}
-  \sum_{\ell=0}^k \p{-\mu}^\ell \epsilon_{qrk}^{(\ell)}
-  Z_{abrs;\mu}^{(\ell)},
-\end{align}
-where $\epsilon_{qrk}^{(\ell)}$ is defined in \eqref{eq:epsilon} and
-\begin{align}
-  Z_{abrs;\mu}^{(\ell)}
-  &\equiv \p{-\mu s+S_z}^\ell \p{\mu\sp{r-s}+S_z}^a S_z^b \\
-  &= \sum_{m=0}^\ell { \ell \choose m } \p{-\mu s}^{\ell-m}
-  \sum_{n=0}^a { a \choose n } \p{\mu\sp{r-s}}^{a-n} S_z^{b+m+n}.
+  \tilde Z_{qrab;\mu}^{(k)}
+  &\equiv \p{\mu\sp{r-k}+S_z}^a
+  Z_{qr;\bmu}^{(k)} \p{\mu\sp{q-k}+S_z}^b \\
+  &= \sum_{\ell=0}^k \p{-1}^\ell \xi_{qrk}^{(\ell)}
+  \sum_{m=0}^a \p{r-k}^{a-m} { a \choose m }
+  \sum_{n=0}^b \p{q-k}^{b-n} { b \choose n }
+  \mu^{a+b+\ell-m-n} S_z^{\ell+m+n}.
 \end{align}
@@
 \begin{align}
-  \sum_j \f12\sp{\sigma_j^\bmu \sigma_j^\mu,\O}_+
-  = \f12\sp{S-\mu S_z,\O}_+
-  = S\O - \f{\mu}{2}\sp{S_z,\O}_+
-  = \O\p{S-\mu S_z} - \f{\mu}{2}\sp{S_z,\O}_-,
+  -\f12 \sp{\sigma_j^\bmu \sigma_j^\mu,\O}_+
+  = -\f12\sp{\p{S-\mu S_z}, \O}_+
+  = -S\O + \f{\mu}{2} \sp{S_z, \O}_+,
 \end{align}
@@
   \D_\mu\p{\O}
-  = \sum_j \sigma_j^\bmu \O \sigma_j^\mu
-  - \O\p{S - \mu S_z} + \f{\mu}{2} \sp{S_z,\O}_-.
+  = \sum_j\sigma_j^\bmu \O \sigma_j^\mu
+    - S \O + \f{\mu}{2} \sp{S_z, \O}_+.
+  \label{eq:D_mu}
 \end{align}
@@
 
-If we consider only ordered collective spin operators, there are two
-cases to work out: the effect of decoherence through $\D_\mu$ on
-$S_\mu^\ell S_\bmu^m S_z^n$, and the effect on
-$S_\bmu^\ell S_\mu^m S_z^n$.  We work out the latter case first,
+If we consider only symmetrically ordered collective spin operators,
+there are two cases to work out: the effect of decoherence through
+$\D_\mu$ on $S_\mu^\ell S_z^m S_\bmu^n$, and the effect on
+$S_\bmu^\ell S_z^m S_\mu^n$.  We work out the latter case first,
 namely
 \begin{align}
-  \D_\mu\p{S_\bmu^\ell S_\mu^m S_z^n}
-  = \sum_j \sigma_j^\bmu S_\bmu^\ell S_\mu^m S_z^n \sigma_j^\mu
-  - S_\bmu^\ell S_\mu^m S_z^n \p{S - \mu S_z}
-  + \f{\mu}{2} \sp{S_z, S_\bmu^\ell S_\mu^m S_z^n}_-.
-  \label{eq:D_mu_nu}
+  \D_\mu\p{S_\bmu^\ell S_z^m S_\mu^n}
+  = \sum_j\sigma_j^\bmu S_\bmu^\ell S_z^m S_\mu^n \sigma_j^\mu
+  - S S_\bmu^\ell S_z^m S_\mu^n
+  + \f{\mu}{2} \sp{S_z, S_\bmu^\ell S_z^m S_\mu^n}_+.
 \end{align}
-The commutator in \eqref{eq:D_mu_nu} is straightforward to simplify
-using \eqref{eq:comm_z_mu}:
+To simplify the first term, we use \eqref{eq:comm_z_mu_base} to push
+the jump operators together, finding
 \begin{align}
-  \mu \sp{S_z, S_\bmu^\ell S_\mu^m S_z^n}_-
-  = \mu \sp{S_z, S_\bmu^\ell S_\mu^m}_- S_z^n
-  = \mu\bmu \p{\ell-m} S_\bmu^\ell S_\mu^m S_z^n
-  = - \p{\ell-m} S_\bmu^\ell S_\mu^m S_z^n,
+  \sum_j \sigma_j^\bmu S_\bmu^\ell S_z^m S_\mu^n \sigma_j^\mu
+  = S_\bmu^\ell \sum_j \sigma_j^\bmu \sigma_j^\mu \p{\mu+S_z}^m S_\mu^n
+  = S_\bmu^\ell \p{S+S_z} \p{\mu+S_z}^m S_\mu^n.
 \end{align}
-while the first term in \eqref{eq:D_mu_nu} can be written in terms of
-collective spin operators if we use \eqref{eq:comm_z_mu_base} to push
-the jump operators together, noting that $\sigma_j^\mu$ commutes with
-$S_\mu$:
+We can likewise push $S_z$ operators together using
+\eqref{eq:push_z_mu_single} to expand
 \begin{align}
-  \sum_j \sigma_j^\bmu S_\bmu^\ell S_\mu^m S_z^n \sigma_j^\mu
-  = \sum_j S_\bmu^\ell \sigma_j^\bmu \sigma_j^\mu S_\mu^m \p{\mu+S_z}^n
-  = S_\bmu^\ell \p{S - \mu S_z} S_\mu^m
-  \sum_{k=0}^n \mu^{n-k} { n \choose k } S_z^k.
-  \label{eq:combine_jumps}
+  \sp{S_z, S_\bmu^\ell S_z^m S_\mu^n}_+
+  = S_z S_\bmu^\ell S_z^m S_\mu^n + S_\bmu^\ell S_z^m S_\mu^n S_z
+  = S_\bmu^\ell\p{\bmu\ell - \mu n + 2 S_z} S_z^m S_\mu^n,
 \end{align}
-Finally, we push the middle $S_z$ to the right using
-\eqref{eq:push_z_mu_single}:
+which implies
 \begin{align}
-  \sum_j \sigma_j^\bmu S_\bmu^\ell S_\mu^m S_z^n \sigma_j^\mu
-  = S_\bmu^\ell S_\mu^m \p{S - m - \mu S_z}
-  \sum_{k=0}^n \mu^{n-k} { n \choose k } S_z^k.
+  \f{\mu}{2} \sp{S_z, S_\bmu^\ell S_z^m S_\mu^n}_+
+  = S_\bmu^\ell\p{-\f12\sp{\ell+n} + \mu S_z} S_z^m S_\mu^n,
 \end{align}
-Altogether, we have that
+and in turn
 \begin{align}
-  \D_\mu\p{S_\bmu^\ell S_\mu^m S_z^n}
-  &= S_\bmu^\ell S_\mu^m \sp{\p{S - m - \mu S_z} \sum_{k=0}^n
-    \mu^{n-k} { n \choose k } S_z^k
-    - \p{S - \mu S_z + \f12\sp{\ell-m}} S_z^n} \\
-  &= S_\bmu^\ell S_\mu^m \sp{\p{S - m - \mu S_z} \sum_{k=0}^{n-1}
-    \mu^{n-k} { n \choose k } S_z^k - \f12 \p{\ell+m} S_z^n}.
+  \D_\mu\p{S_\bmu^\ell S_z^m S_\mu^n}
+  &= S_\bmu^\ell \sp{\p{S+S_z} \p{\mu+S_z}^m
+    - \p{S + \f12\sp{\ell+n} + S_z} S_z^m} S_\mu^n \\
+  &= S_\bmu^\ell
+  \sp{\p{S+S_z} \sum_{k=0}^{m-1} \mu^{m-k} { m \choose k} S_z^k
+  - \f12\p{\ell+n} S_z^m} S_\mu^n.
 \end{align}
 Decoherence through $\D_\mu$ only couples operators of the form
-$S_\bmu^\ell S_\mu^m S_z^n$ to operators of the (almost identical)
-form $S_\bmu^\ell S_\mu^m S_z^{n'}$ with $n'\le n$.
+$S_\bmu^\ell S_z^m S_\mu^n$ to operators of the (almost identical)
+form $S_\bmu^\ell S_z^{m'} S_\mu^n$ with $m'\le m$.
 
@@
 \begin{align}
-  \D_\mu\p{S_\mu^\ell S_\bmu^m S_z^n}
-  = \sum_j \sigma_j^\bmu S_\mu^\ell S_\bmu^m S_z^n \sigma_j^\mu
-  - S_\mu^\ell S_\bmu^m S_z^n \p{S - \mu S_z}
-  + \f{\mu}{2} \sp{S_z, S_\mu^\ell S_\bmu^m S_z^n}_-.
+  \D_\mu\p{S_\mu^\ell S_z^m S_\bmu^n}
+  = \sum_j\sigma_j^\bmu S_\mu^\ell S_z^m S_\bmu^n \sigma_j^\mu
+  - S S_\mu^\ell S_z^m S_\bmu^n
+  + \f{\mu}{2} \sp{S_z, S_\mu^\ell S_z^m S_\bmu^n}_+.
   \label{eq:D_mu_mu}
-\end{align}
-Similarly to before, the commutator in \eqref{eq:D_mu_mu} is
-\begin{align}
-  \mu \sp{S_z, S_\mu^\ell S_\bmu^m S_z^n}_-
-  = \mu \sp{S_z, S_\mu^\ell S_\bmu^m}_- S_z^n
-  = \mu\mu \p{\ell-m} S_\mu^\ell S_\bmu^m S_z^n
-  = \p{\ell-m} S_\mu^\ell S_\bmu^m S_z^n.
 \end{align}
@@
 \begin{align}
-  S_z^n = \sum_{m=-S}^S m^n \op{m},
+  S_z^n = \sum_{k=-S}^S k^n \op{k},
   &&
-  S_\mu = \sum_{m=-S}^S g_\mu\p{m} \op{m+\mu}{m},
+  S_\mu
+  = \sum_{k=-S+\delta_{\mu,-1}}^{S-\delta_{k,1}}
+  g_\mu\p{k} \op{k+\mu}{k},
 \end{align}
@@
 \begin{align}
-  g_\mu\p{S,m} \equiv \sqrt{\p{S-\mu m}\p{S+\mu m+1}},
+  g_\mu\p{S,k} \equiv \sqrt{\p{S-\mu k}\p{S+\mu k+1}},
 \end{align}
-and the initial state
+and the initial states
 \begin{align}
-  \ket{X} \equiv 2^{-N/2} \sum_{m=-S}^S { N \choose S+m }^{1/2} \ket{m},
+  \ket{X} \equiv 2^{-N/2} \sum_{k=-S}^S { N \choose S+k }^{1/2} \ket{k},
+  &&
+  \ket{-Z} \equiv \ket{-S},
 \end{align}
 using which we can compute
-\begin{multline}
-  \bk{X|S_+^\ell S_-^m S_z^n|X}
-  =  2^{-N} \sum_{k=-S}^{S-\max\set{\ell,m}} \p{k+m}^n
-  \sp{{ N \choose S+k+\ell } { N \choose S+k+m }}^{1/2} \\
+\begin{align}
+  \bk{X|S_+^\ell S_z^m S_-^n|X}
+  &=  2^{-N} \sum_{k=-S}^{S-\max\set{\ell,n}} k^m
+  \sp{{ N \choose S+k+\ell } { N \choose S+k+n }}^{1/2} \tag*{} \\
+  &\qquad\qquad\qquad\qquad
   \times \prod_{p=0}^{\ell-1} g_+\p{S,k+p}
-  \prod_{q=0}^{m-1} g_+\p{S,k-q}.
-\end{multline}
+  \prod_{q=0}^{n-1} g_-\p{S,k+n-q} \\
+  &= 2^{-N} \sum_{k=-S}^{S-\max\set{\ell,n}} k^m
+  \f{N!\p{S-k}!}{\p{S+k}!\p{S-k-\ell}!\p{S-k-n}!}.
+\end{align}
+and
+\begin{align}
+  \bk{-Z|S_+^\ell S_z^m S_-^n|-Z}
+  = \delta_{\ell,0} \delta_{n,0} \p{-S}^m.
+\end{align}
 

### 2018-08-28 soc_squeezing/scripts/correlator_methods.py
added -Z expectation value to correlator_methods.py; rearranged some text in expansions notes

@@ -17,7 +17,7 @@ def stirling(n,k): return float(sympy_stirling(n, k, kind = 1, signed = False))
 def transverse_elem(mu, S, m):
     return np.sqrt((S-mu*m)*(S+mu*m+1))
 
-# correlator < X | S_+^l S_z^m S_-^n | X >
+# correlator < +X | S_+^l S_z^m S_-^n | +X >
 def op_val_pX(N, op):
     if op == (0,0,0): return 1
     l, m, n = op
@@ -28,6 +28,12 @@ def op_val_pX(N, op):
         return numerator - denominator
     return sum([ k**m * np.exp( ln_factors(k,l,n) - N*np.log(2) )
                  for k in np.arange(-S,S-max(l,n)+1) ])
+
+# correlator < -Z | S_+^l S_z^m S_-^n | -Z >
+def op_val_nZ(N, op):
+    l, m, n = op
+    if l != 0 or n != 0: return 0
+    return (-N/2)**m
 
 # collective spin operator commutator coefficients (see notes)
 def epsilon(m,n,p,l):

### 2018-08-28 soc_squeezing/cumulant_expansions.tex
added -Z expectation value to correlator_methods.py; rearranged some text in expansions notes
added full, faithful explanation of how time-series correlators are computed, and add derivation of initial values for the state +X

@@
 \newcommand{\D}{\mathcal{D}}
+\newcommand{\F}{\mathcal{F}}
 \newcommand{\I}{\mathcal{I}}
@@
 \end{align}
-where $S\equiv N/2$ and $\D_-\p{S_z^n}$ is worked out in Appendix
-\ref{sec:decoherence}.  In particular,
+where $S\equiv N/2$ throughout this work and $\D_-\p{S_z^n}$ is worked
+out in Appendix \ref{sec:decoherence}.  In particular,
 \begin{align}
@@
 
-\section{Cumulant expansion and Heisenberg systems of equations}
+\section{Cumulant expansion and systems of equations for Heisenberg
+  operators}
 
@@
 is to find equations of motion for operators of interest, and at some
-point truncate high-order terms to arrive at a closed,
-reasonably-sized system of equations.  To benchmark this method, we
-will test it against exact one-axis twisting results.
+point truncate high-order terms with negligible contribution to the
+spin correlators of interest.  To benchmark this method, we test it
+against exact one-axis twisting results in the presence of
+decoherence, as well as transverse-field and two-axis twisting in the
+absence of decoherence.
 
@@
 \begin{align}
-  H = \sum_{p,q,r} h_{pqr} S_+^p S_z^q S_-^r.
+  H = \sum_{p,q,r} h_{pqr} S_+^p S_z^q S_-^r,
   \label{eq:H_general}
 \end{align}
-The Hamiltonians for one-axis twisting with a transverse-field, for
-example, is
+where Hermiticity requires that $h_{pqr}=h_{rqp}$.  The Hamiltonians
+for one-axis twisting with a transverse field, for example, is
 \begin{align}
@@
   + \gamma \D_-\p{S_+^\ell S_z^m S_-^n}
-  \equiv \sum_{p,q,r} T^{\ell m n}_{pqr} S_+^p S_z^q S_-^r,
+  \equiv \sum_{p,q,r} T^{\ell m n}_{pqr} S_+^p S_z^q S_-^r.
 \end{align}
-where the commutator is worked out in Appendix \ref{sec:comm_general},
-and the decoherence operator in Appendix \ref{sec:decoherence}.  We
-can then expand
+Equipped with a matrix representation of the time derivative operator,
+i.e.~$T^{\ell m n}_{pqr}$, we can expand
 \begin{align}
@@
   \bk{\f{d^k}{dt^k}\p{S_+^\ell S_z^m S_-^n}}_{t=0}
-  \approx \sum_{k=0}^{M-1} \f{t^k}{k!}
-  \bk{\f{d^k}{dt^k}\p{S_+^\ell S_z^m S_-^n}}_{t=0},
+  = \sum_{k\ge0} \f{t^k}{k!}
+  \sum_{p,q,r} \sp{T^k}^{\ell m n}_{pqr} \bk{S_+^p S_z^q S_-^r}_{t=0},
+  \label{eq:time_series}
 \end{align}
-where $M$ is chosen such that
+where
 \begin{align}
-  \f{t^M}{M!} \bk{\f{d^M}{dt^M}\p{S_+^\ell S_z^m S_-^n}}_{t=0}
+  \sp{T^k}^{\ell mn}_{pqr}
+  \equiv \sum_{\v\lambda,\v\mu,\v\nu}
+  T^{\ell mn}_{\lambda_1 \mu_1 \nu_1}
+  T^{\lambda_1 \mu_1 \nu_1}_{\lambda_2 \mu_2 \nu_2}
+  T^{\lambda_2 \mu_2 \nu_2}_{\lambda_3 \mu_3 \nu_3}
+  \cdots T^{\lambda_{k-1} \mu_{k-1} \nu_{k-1}}_{pqr}.
+\end{align}
+As terms in \eqref{eq:time_series} are suppressed factorially in $k$,
+for any fixed (or maximal) time $t$ we can truncate the series to
+write
+\begin{align}
+  \bk{S_+^\ell S_z^m S_-^n}
+  \approx \sum_{k=0}^{M_t-1} \f{t^k}{k!}
+  \sum_{p,q,r} \sp{T^k}^{\ell m n}_{pqr} \bk{S_+^p S_z^q S_-^r}_{t=0},
+\end{align}
+where $M_t$ is chosen such that
+\begin{align}
+  \f{t^{M_t}}{M_t!} \sum_{p,q,r}
+  \sp{T^{M_t}}^{\ell m n}_{pqr} \bk{S_+^p S_z^q S_-^r}_{t=0}
   \ll S^{\ell+m+n}.
 \end{align}
-Initial conditions, i.e.~expectation values at $t=0$, are provided in
-Appendix \ref{sec:initial_conditions}.
+The commutator and decoherence operator necessary for computing the
+time derivative operator $T^{\ell m n}_{pqr}$ are respectively worked
+out in Appendix \ref{sec:comm_general} and \ref{sec:decoherence},
+while expectation values for initial conditions of interest are
+provided in Appendix \ref{sec:initial_conditions}.
 % comment on explosion of operators
@@
 \begin{align}
-  S_z^\ell S_\mu^m
-  = S_\mu^m \p{\mu m + S_z}^\ell,
+  S_z^\ell S_\mu^m = S_\mu^m \p{\mu m + S_z}^\ell,
   \label{eq:push_z_mu}
@@
 \begin{align}
+  \ket{-Z} \equiv \ket{-S},
+  &&
   \ket{X} \equiv 2^{-N/2} \sum_{k=-S}^S { N \choose S+k }^{1/2} \ket{k},
-  &&
-  \ket{-Z} \equiv \ket{-S},
 \end{align}
 using which we can compute
+\begin{align}
+  \bk{-Z|S_+^\ell S_z^m S_-^n|-Z}
+  = \delta_{\ell,0} \delta_{m,0} \p{-S}^m,
+\end{align}
+and
 \begin{align}
@@
   \prod_{q=0}^{n-1} g_-\p{S,k+n-q} \\
-  &= 2^{-N} \sum_{k=-S}^{S-\max\set{\ell,n}} k^m
-  \f{N!\p{S-k}!}{\p{S+k}!\p{S-k-\ell}!\p{S-k-n}!}.
+  &= \f{N!}{2^N} \sum_{k=-S}^{S-\max\set{\ell,n}} k^m
+  \f{\p{S-k}!}{\p{S+k}!\p{S-k-\ell}!\p{S-k-n}!}.
+\end{align}
+With the help of Mathematica, we can evaluate the sum to find
+\begin{align}
+  \bk{X|S_+^\ell S_z^m S_-^n|X}
+  = \f{\p{N!}^2}{2^N\p{N-\ell}!\p{N-n}!}
+  \sp{\p{-S}^m \alpha + \f{\p{N-\ell}\p{N-n}}{N} \beta},
+\end{align}
+where $\alpha$ and $\beta$ are defined in terms of the generalized
+hypergeometric function $\F$ by
+\begin{align}
+  \alpha
+  \equiv \F\p{\let\scriptstyle\textstyle
+    \substack{\ell-N,~n-N\\-N}~; -1},
 \end{align}
@@
 \begin{align}
-  \bk{-Z|S_+^\ell S_z^m S_-^n|-Z}
-  = \delta_{\ell,0} \delta_{n,0} \p{-S}^m.
+  \beta
+  \equiv \sum_{p=0}^{m-1} \p{-S}^p { m \choose p }
+  \tilde\F_{m-p-1}\p{\let\scriptstyle\textstyle
+    \substack{1+\ell-N,~1+n-N\\1-N}~; -1},
 \end{align}
+for
+\begin{align}
+  \tilde\F_q\p{\let\scriptstyle\textstyle
+    \substack{a,~b\\c}~; z}
+  \equiv \F\p{\let\scriptstyle\textstyle
+    \substack{2,2,\cdots,2,a,b\\1,1,\cdots,1,c}~; z}
+\end{align}
+with $q$ repetitions of 2 and 1.
 

### 2018-08-29 soc_squeezing/cumulant_expansions.tex
add more thorough explanation of decoherence calculation method

@@
   + \gamma \D_-\p{S_+^\ell S_z^m S_-^n}
-  \equiv \sum_{p,q,r} T^{\ell m n}_{pqr} S_+^p S_z^q S_-^r.
+  \equiv \sum_{p,q,r} T^{\ell m n}_{pqr} S_+^p S_z^q S_-^r,
+  \label{eq:time_derivative}
 \end{align}
-Equipped with a matrix representation of the time derivative operator,
-i.e.~$T^{\ell m n}_{pqr}$, we can expand
+where commutator and decoherence operator necessary for computing the
+matrix elements $T^{\ell m n}_{pqr}$ of the time derivative operator
+$d/dt$ are respectively worked out in Appendix \ref{sec:comm_general}
+and \ref{sec:decoherence}.
+
+The time derivative operator $d/dt$ will generally couple spin
+operators $S_+^\ell S_z^m S_-^n$ to spin operators $S_+^p S_z^q S_-^r$
+with higher weight, i.e.~with $p+q+r>\ell+m+n$.  In practice, keeping
+track of this operator growth will eventually require more
+computational resources than are available, so we must truncate it at
+some point by taking
+\begin{align}
+  \f{d}{dt} \p{S_+^\ell S_z^m S_-^n}
+  = \sum_{p,q,r} T^{\ell m n}_{pqr} S_+^p S_z^q S_-^r
+  \to \sum_{w\p{p,q,r}<W} T^{\ell m n}_{pqr} S_+^p S_z^q S_-^r
+  \label{eq:time_derivative_truncation}
+\end{align}
+for some cutoff $W$ and weight function $w$, e.g.~$w\p{p,q,r}=p+q+r$,
+although the most optimal weight function will generally be
+system-dependent.  The truncation in
+\eqref{eq:time_derivative_truncation} will close the system of
+differential equations defined by \eqref{eq:time_derivative}, and
+allow us to solve it using standard numerical methods.  The
+expectation values of collective spin operators for initial conditions
+of interest are provided in Appendix \ref{sec:initial_conditions}.
+
+Alternately, we can formally solve the differential equation in
+\eqref{eq:time_derivative} and write
+\begin{align}
+  S_+^\ell S_z^m S_-^n
+  = \left.\exp\p{t \f{d}{dt}}\right|_{t=0} \p{S_+^\ell S_z^m S_-^n}
+  = \sum_{k\ge0} \f{t^k}{k!}
+  \left.\f{d^k}{dt^k}\right|_{t=0} \p{S_+^\ell S_z^m S_-^n},
+\end{align}
+which implies
 \begin{align}
@@
   = \sum_{k\ge0} \f{t^k}{k!}
-  \bk{\f{d^k}{dt^k}\p{S_+^\ell S_z^m S_-^n}}_{t=0}
-  = \sum_{k\ge0} \f{t^k}{k!}
-  \sum_{p,q,r} \sp{T^k}^{\ell m n}_{pqr} \bk{S_+^p S_z^q S_-^r}_{t=0},
+  \sum_{p,q,r} T^{\ell m n}_{pqr;k} \bk{S_+^p S_z^q S_-^r}_{t=0},
   \label{eq:time_series}
 \end{align}
-where
+where the matrix elements of the $k$-th time derivative are given by
 \begin{align}
-  \sp{T^k}^{\ell mn}_{pqr}
+  T^{\ell m n}_{pqr;0}=\delta_{\ell p}\delta_{mq}\delta_{nr},
+  &&
+  T^{\ell m n}_{pqr;1}\equiv T^{\ell m n}_{pqr},
+\end{align}
+\begin{align}
+  T^{\ell mn}_{pqr;k>1}
   \equiv \sum_{\v\lambda,\v\mu,\v\nu}
@@
 \end{align}
-As terms in \eqref{eq:time_series} are suppressed factorially in $k$,
-for any fixed (or maximal) time $t$ we can truncate the series to
-write
+Instead of explicitly choosing a weight function that determines which
+operators to keep track of, we can now truncate the series in
+\eqref{eq:time_series} by taking
 \begin{align}
   \bk{S_+^\ell S_z^m S_-^n}
-  \approx \sum_{k=0}^{M_t-1} \f{t^k}{k!}
-  \sum_{p,q,r} \sp{T^k}^{\ell m n}_{pqr} \bk{S_+^p S_z^q S_-^r}_{t=0},
+  = \sum_{k\ge0} \f{t^k}{k!}
+  \sum_{p,q,r} T^{\ell m n}_{pqr;k} \bk{S_+^p S_z^q S_-^r}_{t=0}
+  \to \sum_{k<M} \f{t^k}{k!}
+  \sum_{p,q,r} T^{\ell m n}_{pqr;k} \bk{S_+^p S_z^q S_-^r}_{t=0}.
 \end{align}
-where $M_t$ is chosen such that
-\begin{align}
-  \f{t^{M_t}}{M_t!} \sum_{p,q,r}
-  \sp{T^{M_t}}^{\ell m n}_{pqr} \bk{S_+^p S_z^q S_-^r}_{t=0}
-  \ll S^{\ell+m+n}.
-\end{align}
-The commutator and decoherence operator necessary for computing the
-time derivative operator $T^{\ell m n}_{pqr}$ are respectively worked
-out in Appendix \ref{sec:comm_general} and \ref{sec:decoherence},
-while expectation values for initial conditions of interest are
-provided in Appendix \ref{sec:initial_conditions}.
-% comment on explosion of operators
+Examining the nonzero entries of $T^{\ell mn}_{pqr;k}$ for
+$k=1,2,\cdots,M-1$ now {\it tells} us which operators we need to keep
+track of in order to compute the expectation value
+$\bk{S_+^\ell S_z^m S_-^n}$.  Furthermore, after computing the
+expectation values of all relevant operators at $t=0$, time evolution
+of $\bk{S_+^\ell S_z^m S_-^n}$ takes only $\O\p{M}$ CPU time.
+
+As a final comment, we note that the relation
+$\p{S_+^\ell S_z^m S_-^n}^\dag=S_+^n S_z^m S_-^\ell$ can be exploited
+to cut the number of operators we need to keep track of roughly in
+two.
 
@@
   &= \mu S_\mu,
-  \label{eq:comm_z_mu_base} \\
+  \label{eq:comm_z_base} \\
   \sp{\sigma_j^\mu,\sigma_k^\bmu}_-
@@
 \end{align}
-These relations can be used to inductively determine commutation
-relations with higher-weight spin operators.  For example, we can find
-that
+These relations can be used to inductively compute identities
+involving powers of collective spin operators.  Working through the
+possibilities, we find
 \begin{align}
-  S_z S_\mu^m
-  = \p{\mu S_\mu + S_\mu S_z} S_\mu^{m-1}
-  = \mu S_\mu^m + S_\mu S_z S_\mu^{m-1}
-  = 2 \mu S_\mu^m + S_\mu S_z S_\mu^{m-2}
-  = S_\mu^m \p{\mu m + S_z}.
+  S_z^m \sigma_j^\mu
+  = S_z^{m-1} \sigma_j^\mu \p{\mu + S_z}
+  = S_z^{m-2} \sigma_j^\mu \p{\mu + S_z}^2
+  = \cdots
+  = \sigma_j^\mu \p{\mu + S_z}^m,
+  \label{eq:push_z_mu_Ss}
+\end{align}
+and
+\begin{align}
+  \sigma_j^z S_\mu^m
+  = \p{2\mu\sigma_j^\mu + S_\mu \sigma_j^z} S_\mu^{m-1}
+  = 2\mu\sigma_j^\mu S_\mu^{m-1} + S_\mu \sigma_j^z S_\mu^{m-1}
+  = \cdots
+  = 2\mu m\sigma_j^\mu S_\mu^{m-1} + S_\mu^m \sigma_j^z.
+  \label{eq:push_z_mu_sS}
+\end{align}
+Summing over the single-spin index $j$ in both of these cases gives us
+the purely collective-spin versions of these identities:
+\begin{align}
+  S_z^m S_\mu = S_\mu \p{\mu + S_z}^m,
+  &&
+  S_z S_\mu^m = S_\mu^m \p{\mu m + S_z},
   \label{eq:push_z_mu_single}
 \end{align}
-Repeating this process $\ell$ times tells us
+where we can repeat the process of pushing through individual
+operators $\ell$ times to get
 \begin{align}
-  S_z^\ell S_\mu^m = S_\mu^m \p{\mu m + S_z}^\ell,
+  S_z^\ell S_\mu^m
+  = S_z^{\ell-1} S_\mu^m \p{\mu m + S_z}
+  = S_z^{\ell-2} S_\mu^m \p{\mu m + S_z}^2
+  = \cdots
+  = S_\mu^m \p{\mu m + S_z}^\ell.
   \label{eq:push_z_mu}
 \end{align}
-which implies
-\begin{align}
-  \sp{S_z^\ell, S_\mu^m}_-
-  = S_\mu^m \sp{\p{\mu m + S_z}^\ell - S_z^\ell}
-  = S_\mu^m \sum_{k=0}^{\ell-1} {\ell \choose k} \p{\mu m}^{\ell-k} S_z^k.
-  \label{eq:comm_z_mu}
-\end{align}
-Similarly finding commutation relations between high-weight transverse
-spin operators, i.e.~$S_\mu^m$ and $S_\bmu^n$, turns out to be
-considerably more difficult than the cases we have worked out thus
-far.  We therefore save this work for Appendix
-\ref{sec:comm_transverse}.
+Similarly finding commutation relations between powers of transverse
+spin operators, i.e.~$S_\mu$, turns out to be considerably more
+difficult than the cases we have worked out thus far.  We therefore
+save this work for Appendix \ref{sec:comm_transverse}.
 
@@
 
-To determine commutation relations between high-weight collective spin
-operators transverse to the decoherence (i.e.~$z$) axis, we compute
-(with $\bmu\equiv-\mu\in\set{+1,-1}$)
+To find commutation relations between powers of transverse collective
+spin operators, we compute (with $\bmu\equiv-\mu\in\set{+1,-1}$)
 \begin{align}
-  S_\mu S_\bmu^n
-  &= S_\bmu S_\mu S_\bmu^{n-1} + 2\mu S_z S_\bmu^{n-1} \\
-  &= S_\bmu \sp{S_\bmu S_\mu S_\bmu^{n-2} + 2\mu S_z S_\bmu^{n-2}}
-  + 2\mu S_z S_\bmu^{n-1} \\
-  &= S_\bmu^2 S_\mu S_\bmu^{n-2}
-  + 2\mu \p{S_\bmu S_z S_\bmu ^{n-2} + S_z S_\bmu^{n-1}} \\
-  &= S_\bmu^n S_\mu + 2\mu \sum_{k=0}^{n-1} S_\bmu^k S_z S_\bmu^{n-k-1}.
-  \label{eq:comm_mu_single}
+  S_\mu^m \sigma_j^\bmu
+  &= S_\mu^{m-1}\sigma_j^\bmu S_\mu + \mu S_\mu^{m-1} \sigma_j^z \\
+  &= S_\mu^{m-2} \sigma_j^\bmu S_\mu^2 + \mu S_\mu^{m-2} \sigma_j^z S_\mu
+  + \mu S_\mu^{m-1} \sigma_j^z \\
+  &= \sigma_j^\bmu S_\mu^m
+  + \mu \sum_{k=0}^{m-1} S_\mu^k \sigma_j^z S_\mu^{m-k-1}
+  \label{eq:push_mu_Ss_start}.
 \end{align}
-While \eqref{eq:comm_mu_single} gives us the commutator
-$\sp{S_\mu,S_\bmu^n}_-$, we would like to enforce an ordering on
-products of collective spin operators, which allows us to only keep
-track of high-weight operators that are linearly independent.  We
-choose to impose an ordering for which all collective spin operators
-take the form $S_\bmu^\ell S_\mu^m S_z^n$, as such an ordering will
-prove convenient for the calculations in this section.  This choice of
+While \eqref{eq:push_mu_Ss_start} gives us the commutator
+$\sp{S_\mu^m,\sigma_j^\bmu}_-$, we would like to enforce an ordering
+on products of spin operators, which will allow us to only keep track
+of operators that are linearly independent.  We choose to impose an
+ordering with all $\sigma_j^\bmu$ operators on the left, and all
+$\sigma_j^z$ operators on the right, as such an ordering will prove
+convenient for the calculations in this section.  This choice of
 ordering compels us to expand
 \begin{align}
-  2 \mu \sum_{k=0}^{n-1} S_\bmu^k S_z S_\bmu^{n-k-1}
-  = 2 \mu \sum_{k=0}^{n-1}
-  S_\bmu^k S_\bmu^{n-k-1} \p{\bmu\sp{n-k-1} + S_z}
-  = n S_\bmu^{n-1}\p{1 - n + 2\mu S_z},
+  \mu \sum_{k=0}^{m-1} S_\mu^k \sigma_j^z S_\mu^{m-k-1}
+  &= \mu \sum_{k=0}^{mq-1} S_\mu^k
+  \p{2\mu\sp{m-k-1} \sigma_j^\mu S_\mu^{m-k-2}
+    + S_\mu^{m-k-1} \sigma_j^z} \\
+  &= m \p{m-1} \sigma_j^\mu S_\mu^{m-2} + \mu m S_\mu^{m-1} \sigma_j^z,
 \end{align}
-so
+which implies
 \begin{align}
-  \sp{S_\mu, S_\bmu^n}_-
-  = n S_\bmu^{n-1}\p{1 - n + 2\mu S_z}.
+  S_\mu^m \sigma_j^\bmu
+  = \sigma_j^\bmu S_\mu^m + m \p{m-1} \sigma_j^\mu S_\mu^{m-2}
+  + \mu m S_\mu^{m-1} \sigma_j^z,
+  \label{eq:push_mu_Ss}
+\end{align}
+and in turn
+\begin{align}
+  S_\mu^m S_\bmu = S_\bmu S_\mu^m
+  + m S_\mu^{m-1} \p{m - 1 + 2\mu S_z}.
+  \label{eq:push_mu_single}
 \end{align}
@@
 where in this final form $E_{mn;\mu}^{(0)} = D_{mn;\mu}$, which
-together with \eqref{eq:C_mn_E} implies that
+together with the expansion for $C_{mn;\mu}$ in \eqref{eq:C_mn_E}
+implies that
 \begin{align}
@@
 
-\section{Decoherence of collective spin operators through individual
+\section{Decoherence of collective spin operators from uncorrelated
   spin decay}
@@
 \end{align}
-To simplify the first term, we use \eqref{eq:comm_z_mu_base} to push
-the jump operators together, finding
+To simplify the first term, we use \eqref{eq:comm_z_base} to push the
+jump operators together, finding
 \begin{align}
@@
 \begin{align}
-  \ket{-Z} \equiv \ket{-S},
+  \ket{\pm Z} \equiv \ket{\pm S},
   &&
@@
 \begin{align}
-  \bk{-Z|S_+^\ell S_z^m S_-^n|-Z}
-  = \delta_{\ell,0} \delta_{m,0} \p{-S}^m,
+  \bk{\pm Z|S_+^\ell S_z^m S_-^n|\pm Z}
+  = \delta_{\ell,0} \delta_{m,0} \p{\pm S}^m,
 \end{align}
@@
 \end{align}
-With the help of Mathematica, we can evaluate the sum to find
+With the help of Mathematica, we can write the sum in the alternate
+form
 \begin{align}

### 2018-09-04 soc_squeezing/cumulant_expansions.tex
improved text in main body and added all decoherence calculations

@@
 \newcommand{\I}{\mathcal{I}}
+\newcommand{\J}{\mathcal{J}}
 \renewcommand{\O}{\mathcal{O}}
@@
 \newcommand{\N}{\mathbb{N}}
-\newcommand{\Z}{\mathbb{Z}}
-
+
+\newcommand{\z}{\text{z}}
+\newcommand{\x}{\text{x}}
+\newcommand{\y}{\text{y}}
+\newcommand{\Z}{\text{Z}}
+\newcommand{\X}{\text{X}}
 \newcommand{\bmu}{{\bar\mu}}
@@
 
-Previous work has shown how to realize various spin-squeezing
-Hamiltonians with ultracold atomic systems.  While it is
-straightforward to simulate the dynamics generated by these
-Hamiltonians in the context of a closed quantum system, here we try to
-develop an understanding of their the squeezing behavior in the
-presence of decoherence.  In particular, we will be considering
-dynamics induced by the collective one-axis twisting (OAT), two-axis
-twisting (TAT), and transverse field (TVF) Hamiltonians of the form
-\begin{align}
-  H_{\t{OAT}} = \chi S_z^2,
-  &&
-  H_{\t{TAT}}^{z,y} = \chi \p{S_z^2 - S_y^2},
-  &&
-  H_{\t{TVF}} = \chi S_z^2 + \Omega S_x,
-\end{align}
-for spins which decay in an uncorrelated manner at a rate $\gamma$.
-The evolution of any Heisenburg operator $\O$ with a Hamiltonian $H$
-is given by
-\begin{align}
-  \f{d}{dt} \O
-  = i\sp{H,\O}_- + \gamma\D_-\p{\O},
-  &&
-  \D_-\p{\O} \equiv \sum_j\p{\sigma_j^+\O\sigma_j^-
-    - \f12\sp{\sigma_j^+\sigma_j^-,\O}_+},
+\tableofcontents
+
+\newpage
+
+\section{Introduction}
+
+Past work has shown how to realize spin-squeezing Hamiltonians with a
+variety of ultracold atomic and trapped ion systems.  While simulating
+the dynamics and squeezing behavior induced by these Hamiltonians is
+straightforward in the context of a closed quantum system, here we
+develop a method to simulate dynamics in the presence of decoherence.
+In particular, we will consider dynamics induced by the one-axis
+twisting (OAT), two-axis twisting (TAT), and transverse field (TVF)
+Hamiltonians of the form
+\begin{align}
+  H_{\t{OAT}} = \chi S_\z^2,
+  &&
+  H_{\t{TAT}}^{\z,\y} = \chi \p{S_\z^2 - S_\y^2},
+  &&
+  H_{\t{TVF}} = \chi S_\z^2 + \Omega S_\x.
+\end{align}
+In order to determine the squeezing behavior of these Hamiltonians, we
+will study the dynamics of Heisenburg operators $\O$, whose evolution
+under a Hamiltonian $H$ is given by
+\begin{align}
+  \f{d}{dt} \O = i\sp{H,\O}_- + \sum_\J \gamma_\J \D\p{\J;\O},
+  &&
+  \D\p{\J;\O}
+  \equiv \sum_{J\in\J}\p{J^\dag \O J - \f12\sp{J^\dag J,\O}_+},
   \label{eq:EOM}
 \end{align}
-where $\sp{A,B}_\pm\equiv AB\pm BA$ and $j$ indexes an individual
-spin.  Our measure of squeezing is
+where $\J$ is a set of jump operators with a corresponding decoherence
+rate $\gamma_\J$ and $\sp{A,B}_\pm\equiv AB\pm BA$.  Our measure of
+spin squeezing is
 \begin{align}
@@
 \begin{align}
-  H_{\t{OAT}} = \f14 \chi \sum_{j,k} \sigma_j^z \sigma_k^z
-  = \f12 \chi \sum_{j<k} \sigma_j^z \sigma_k^z + \f14 N \chi,
+  H_{\t{OAT}} = \f14 \chi \sum_{j,k} \sigma_\z^{(j)} \sigma_\z^{(k)}
+  = \f12 \chi \sum_{j<k} \sigma_\z^{(j)} \sigma_\z^{(k)} + \f14 N \chi,
 \end{align}
@@
 al.~in ref.~\cite{foss-feig2013nonequilibrium}.  Adapting exact
-expectation values with respect to the initial state $\ket{X}$
-satisfying $S_x\ket{X}=S\ket{X}$, in terms of $\mu,\nu\in\set{+1,-1}$
-we find that
+expectation values with respect to the initial state $\ket{\X}$
+satisfying $S_\x\ket{\X}=S\ket{\X}$, in terms of
+$\mu,\nu\in\set{+1,-1}$ we find that
 \begin{align}
   \bk{S_+}
-  &= \sum_j\bk{\sigma_j^+}
+  &= \sum_j\bk{\sigma_+^{(j)}}
   = \f{N}{2} e^{-\gamma t/2} \Phi\p{\chi,t}^{N-1}, \label{eq:S+_OAT} \\
-  \bk{S_\mu S_z}
-  &= \f12\sum_j\bk{\sigma_j^\mu\sigma_j^z}
-  + \f12\sum_{j\ne k} \bk{\sigma_j^\mu \sigma_k^z} \\
+  \bk{S_\mu S_\z}
+  &= \f12\sum_j\bk{\sigma_\mu^{(j)}\sigma_\z^{(j)}}
+  + \f12\sum_{j\ne k} \bk{\sigma_\mu^{(j)} \sigma_\z^{(k)}} \\
   &= -\f{\mu}{2}\bk{S_\mu} + \f14 N \p{N-1} e^{-\gamma t/2}
@@
   \bk{S_\mu S_\nu}
-  &= \sum_j \bk{\sigma_j^\mu \sigma_j^\nu}
-  + \sum_{j\ne k} \bk{\sigma_j^\mu \sigma_k^\nu} \\
-  &= \delta_{\mu,-\nu} \p{\f{N}{2} + \mu \bk{S_z}}
+  &= \sum_j \bk{\sigma_\mu^{(j)} \sigma_\nu^{(j)}}
+  + \sum_{j\ne k} \bk{\sigma_\mu^{(j)} \sigma_\nu^{(k)}} \\
+  &= \delta_{\mu,-\nu} \p{\f{N}{2} + \mu \bk{S_\z}}
   + \f14 N \p{N-1} e^{-\gamma t}
@@
 \begin{align}
-  \Phi\p{\xi,t} \equiv e^{-\gamma t/2}
-  \sp{\cos\p{s_\xi t}+\f12\gamma t~\t{sinc}\p{s_\xi t}},
-  &&
-  \Psi\p{\xi,t} \equiv e^{-\gamma t/2}
-  \p{is_\xi-\gamma/2}t~ \t{sinc}\p{s_\xi t},
-\end{align}
-and $s_\xi \equiv \xi + i2\gamma$.  The last ingredients we need
-to compute spin squeezing at any time are $\bk{S_z}$ and $\bk{S_z^2}$.
-As these operators commute with the OAT Hamiltonian, their evolution
-is governed entirely by decoherence, such that
-\begin{align}
-  \f{d}{dt}\p{S_z^n}
-  = \gamma \D_-\p{S_z^n}
-  = \gamma \p{S + S_z} \sum_{k=0}^{n-1}
-  \p{-1}^{n-k} { n \choose k } S_z^k,
-\end{align}
-where $S\equiv N/2$ throughout this work and $\D_-\p{S_z^n}$ is worked
-out in Appendix \ref{sec:decoherence}.  In particular,
-\begin{align}
-  \f{d}{dt} S_z = -\gamma \p{S + S_z},
-  &&
-  \f{d}{dt}\p{S_z^2} = \gamma \p{S + S_z} \p{1 - 2S_z}.
-\end{align}
-The initial conditions $\bk{S_z}_{t=0}=0$ and $\bk{S_z^2}_{t=0}=S/2$
-then imply
-\begin{align}
-  \bk{S_z} = S\p{e^{-\gamma t}-1},
-  &&
-  \bk{S_z^2} = S \p{1 - \f12 e^{-\gamma t}} e^{-\gamma t} + \bk{S_z}^2.
+  \Phi\p{\zeta,t} \equiv e^{-\gamma t/2}
+  \sp{\cos\p{s_\zeta t}+\f12\gamma t~\t{sinc}\p{s_\zeta t}},
+  &&
+  \Psi\p{\zeta,t} \equiv e^{-\gamma t/2}
+  \p{is_\zeta-\gamma/2}t~ \t{sinc}\p{s_\zeta t},
+\end{align}
+and $s_\zeta \equiv \zeta + i2\gamma$.  The last ingredients we need
+to compute spin squeezing at any time are $\bk{S_\z}$ and
+$\bk{S_\z^2}$.  As these operators commute with the OAT Hamiltonian,
+their evolution is governed entirely by decoherence through
+uncorrelated decay of individual spins, which means (see Appendix
+\ref{sec:decay_single})
+\begin{align}
+  \f{d}{dt} S_\z = -\gamma \p{S + S_\z},
+  &&
+  \f{d}{dt}\p{S_\z^2} = \gamma \p{S + S_\z} \p{1 - 2S_\z},
+\end{align}
+where $S\equiv N/2$.  The initial conditions $\bk{S_\z}_{t=0}=0$ and
+$\bk{S_\z^2}_{t=0}=S/2$ then imply
+\begin{align}
+  \bk{S_\z} = S\p{e^{-\gamma t}-1},
+  &&
+  \bk{S_\z^2} = S \p{1 - \f12 e^{-\gamma t}} e^{-\gamma t} + \bk{S_\z}^2.
   \label{eq:Sz_OAT}
@@
 \eqref{eq:Sz_OAT} are sufficient to compute the squeezing parameter
-$\xi^2$ in \eqref{eq:squeezing} at any time throughout evolution of
-the initial state $\ket{X}:S_x\ket{X}=S\ket{X}$ under $H_{\t{OAT}}$.
-
-\section{Cumulant expansion and systems of equations for Heisenberg
-  operators}
+$\xi^2$ defined in \eqref{eq:squeezing} at any time throughout
+evolution of the initial state $\ket{\X}:S_\x\ket{\X}=S\ket{\X}$ under
+$H_{\t{OAT}}$.
+
+\section{Truncated Heisenberg operator dynamics}
 
@@
 solutions. We will therefore solve these Hamiltonians approximately
-using systems of equations for Heisenberg operators.  The basic idea
-is to find equations of motion for operators of interest, and at some
-point truncate high-order terms with negligible contribution to the
-spin correlators of interest.  To benchmark this method, we test it
-against exact one-axis twisting results in the presence of
-decoherence, as well as transverse-field and two-axis twisting in the
-absence of decoherence.
+using alternate methods.  The basic strategy will be to find equations
+of motion for operators of interest, and at some point truncate
+high-order terms with negligible contribution to spin correlators of
+interest.  To benchmark this method, we test it against exact one-axis
+twisting results in the presence of decoherence, collective
+Hamiltonian dynamics in the absence of decoherence, and exact
+simulations of few-spin systems.
 
@@
 \begin{align}
-  H = \sum_{p,q,r} h_{pqr} S_+^p S_z^q S_-^r,
+  H = \sum_{p,q,r} h_{pqr} S_+^p S_\z^q S_-^r,
   \label{eq:H_general}
 \end{align}
-where Hermiticity requires that $h_{pqr}=h_{rqp}$.  The Hamiltonians
+where Hermiticity requires that $h_{pqr}=h_{rqp}$.  The Hamiltonian
 for one-axis twisting with a transverse field, for example, is
@@
   H_{\t{TVF}}
-  = \chi S_z^2 + \Omega S_x
-  = \chi S_z^2 + \f{\Omega}{2} S_+ + \f{\Omega}{2} S_-,
-\end{align}
-while the two-axis twisting Hamiltonians about $\p{z,y}$ and $\p{y,x}$
-are
-\begin{align}
-  H_{\t{TAT}}^{z,y}
-  = \chi \p{S_z^2 - S_y^2}
-  = \chi \p{S_z^2 + \f14 S_+^2 + \f14 S_-^2 - \f12 S_+ S_- + \f12 S_z},
+  = \chi S_\z^2 + \Omega S_\x
+  = \chi S_\z^2 + \f{\Omega}{2} S_+ + \f{\Omega}{2} S_-,
+\end{align}
+while the two-axis twisting Hamiltonians about $\p{\z,\y}$ and
+$\p{\y,\x}$ are
+\begin{align}
+  H_{\t{TAT}}^{\z,\y}
+  \propto S_\z^2 - S_\y^2
+  = S_\z^2 + \f14 S_+^2 + \f14 S_-^2
+  - \f12 S_+ S_- + \f12 S_\z,
 \end{align}
@@
 \begin{align}
-  H_{\t{TAT}}^{y,x}
-  = \chi\p{S_y^2 - S_x^2}
-  = -\f{\chi}{2} \p{S_+^2 + S_-^2}.
-\end{align}
-The evolution of a general operator $S_+^\ell S_-^m S_z^n$ under a
+  H_{\t{TAT}}^{\y,\x}
+  \propto S_\y^2 - S_\x^2
+  = -\f12 S_+^2 - \f12 S_-^2.
+\end{align}
+The evolution of a general operator $S_+^\ell S_-^m S_\z^n$ under a
 Hamiltonian of the form in \eqref{eq:H_general} is given by
 \begin{align}
-  \f{d}{dt} \p{S_+^\ell S_z^m S_-^n}
-  = i\sum_{p,q,r} h_{pqr} \sp{S_+^p S_z^q S_-^r, S_+^\ell S_z^m S_-^n}_-
-  + \gamma \D_-\p{S_+^\ell S_z^m S_-^n}
-  \equiv \sum_{p,q,r} T^{\ell m n}_{pqr} S_+^p S_z^q S_-^r,
+  \f{d}{dt} \p{S_+^\ell S_\z^m S_-^n}
+  = i \sum_{p,q,r} h_{pqr}
+  \sp{S_+^p S_\z^q S_-^r, S_+^\ell S_\z^m S_-^n}_-
+  + \sum_\J \gamma_\J \D\p{\J;\O}
+  \equiv \sum_{p,q,r} T^{\ell m n}_{pqr} S_+^p S_\z^q S_-^r,
   \label{eq:time_derivative}
 \end{align}
-where commutator and decoherence operator necessary for computing the
+where commutator and decoherence terms necessary for computing the
 matrix elements $T^{\ell m n}_{pqr}$ of the time derivative operator
-$d/dt$ are respectively worked out in Appendix \ref{sec:comm_general}
-and \ref{sec:decoherence}.
+$d/dt$ are worked out Appendices \ref{sec:comm_general},
+\ref{sec:decoherence_single}, and \ref{sec:decoherence_collective}.
 
 The time derivative operator $d/dt$ will generally couple spin
-operators $S_+^\ell S_z^m S_-^n$ to spin operators $S_+^p S_z^q S_-^r$
-with higher weight, i.e.~with $p+q+r>\ell+m+n$.  In practice, keeping
-track of this operator growth will eventually require more
-computational resources than are available, so we must truncate it at
-some point by taking
-\begin{align}
-  \f{d}{dt} \p{S_+^\ell S_z^m S_-^n}
-  = \sum_{p,q,r} T^{\ell m n}_{pqr} S_+^p S_z^q S_-^r
-  \to \sum_{w\p{p,q,r}<W} T^{\ell m n}_{pqr} S_+^p S_z^q S_-^r
+operators $S_+^\ell S_\z^m S_-^n$ to spin operators
+$S_+^p S_\z^q S_-^r$ with higher weight, i.e.~with $p+q+r>\ell+m+n$.
+In practice, keeping track of this operator growth will eventually
+require more computational resources than are available, so we must
+truncate it at some point by taking
+\begin{align}
+  \f{d}{dt} \p{S_+^\ell S_\z^m S_-^n}
+  = \sum_{p,q,r} T^{\ell m n}_{pqr} S_+^p S_\z^q S_-^r
+  \to \sum_{w\p{p,q,r}<W} T^{\ell m n}_{pqr} S_+^p S_\z^q S_-^r
   \label{eq:time_derivative_truncation}
 \end{align}
-for some cutoff $W$ and weight function $w$, e.g.~$w\p{p,q,r}=p+q+r$,
-although the most optimal weight function will generally be
-system-dependent.  The truncation in
-\eqref{eq:time_derivative_truncation} will close the system of
-differential equations defined by \eqref{eq:time_derivative}, and
-allow us to solve it using standard numerical methods.  The
-expectation values of collective spin operators for initial conditions
-of interest are provided in Appendix \ref{sec:initial_conditions}.
-
-Alternately, we can formally solve the differential equation in
-\eqref{eq:time_derivative} and write
-\begin{align}
-  S_+^\ell S_z^m S_-^n
-  = \left.\exp\p{t \f{d}{dt}}\right|_{t=0} \p{S_+^\ell S_z^m S_-^n}
+for some cutoff $W$ and weight function $w$, e.g.~$w\p{p,q,r}=p+q+r$.
+The truncation in \eqref{eq:time_derivative_truncation} will close the
+system of differential equations defined by
+\eqref{eq:time_derivative}, and allow us to solve it using standard
+numerical methods.  The initial conditions for this system of
+differential equations, i.e.~expectation values of operators with
+respect to states that are simple to prepare, are provided in Appendix
+\ref{sec:initial_conditions}.
+
+The truncation strategy in \eqref{eq:time_derivative_truncation} has a
+few limitations:
+\begin{enumerate*}[label=(\roman*)]
+\item simulating a system of differential equations for a large number
+  of operators can be very time-consuming,
+\item the weight function $w$ must be chosen carefully, as the optimal
+  weight function is generally system-dependent,
+\item simulation results can only be trusted up to the time at which
+  the initial value of operators $S_+^p S_\z^r S_-^q$ with weight
+  $w\p{p,q,r}\ge W$ have a non-negligible contribution to operators of
+  interest.
+  \label{pt:limitation}
+\end{enumerate*}
+While the first two of these limitations might in principle be avoided
+by alternate means of tracking operators, the last limitation
+necessarily applies in some form to any method tracking only a subset
+of {\it all} operators.  Embracing the unavoidable, we therefore try
+to devise an alternate truncation strategy which is built around
+\ref{pt:limitation}.
+
+We can expand operators in a Taylor series about $t=0$ to write
+\begin{align}
+  S_+^\ell S_\z^m S_-^n
   = \sum_{k\ge0} \f{t^k}{k!}
-  \left.\f{d^k}{dt^k}\right|_{t=0} \p{S_+^\ell S_z^m S_-^n},
+  \left.\f{d^k}{dt^k}\right|_{t=0} \p{S_+^\ell S_\z^m S_-^n},
 \end{align}
@@
 \begin{align}
-  \bk{S_+^\ell S_z^m S_-^n}
+  \bk{S_+^\ell S_\z^m S_-^n}
   = \sum_{k\ge0} \f{t^k}{k!}
-  \sum_{p,q,r} T^{\ell m n}_{pqr;k} \bk{S_+^p S_z^q S_-^r}_{t=0},
+  \sum_{p,q,r} T^{\ell m n}_{pqr;k} \bk{S_+^p S_\z^q S_-^r}_{t=0},
   \label{eq:time_series}
@@
 \begin{align}
-  \bk{S_+^\ell S_z^m S_-^n}
+  \bk{S_+^\ell S_\z^m S_-^n}
   = \sum_{k\ge0} \f{t^k}{k!}
-  \sum_{p,q,r} T^{\ell m n}_{pqr;k} \bk{S_+^p S_z^q S_-^r}_{t=0}
+  \sum_{p,q,r} T^{\ell m n}_{pqr;k} \bk{S_+^p S_\z^q S_-^r}_{t=0}
   \to \sum_{k<M} \f{t^k}{k!}
-  \sum_{p,q,r} T^{\ell m n}_{pqr;k} \bk{S_+^p S_z^q S_-^r}_{t=0}.
+  \sum_{p,q,r} T^{\ell m n}_{pqr;k} \bk{S_+^p S_\z^q S_-^r}_{t=0}.
 \end{align}
@@
 track of in order to compute the expectation value
-$\bk{S_+^\ell S_z^m S_-^n}$.  Furthermore, after computing the
+$\bk{S_+^\ell S_\z^m S_-^n}$.  Furthermore, after computing the
 expectation values of all relevant operators at $t=0$, time evolution
-of $\bk{S_+^\ell S_z^m S_-^n}$ takes only $\O\p{M}$ CPU time.
-
-As a final comment, we note that the relation
-$\p{S_+^\ell S_z^m S_-^n}^\dag=S_+^n S_z^m S_-^\ell$ can be exploited
-to cut the number of operators we need to keep track of roughly in
-two.
+of $\bk{S_+^\ell S_\z^m S_-^n}$ takes only $\O\p{M}$ CPU time.
+Finally, the relation
+$\p{S_+^\ell S_\z^m S_-^n}^\dag=S_+^n S_\z^m S_-^\ell$ can be
+exploited to cut both the number of expectation values
+$\bk{S_+^p S_\z^r S_-^q}_{t=0}$ and the number of matrix elements
+$T^{\ell mn}_{pqr}$ which we need to compute roughly in half.
 
@@
 
-This text makes ubiquitous use of various spin operator identities; we
-collect and derive particularly useful identities here for reference.
-The elementary commutation relations between single-weight spin
+The appendices in this text make ubiquitous use of various spin
+operator identities; we collect and derive some basic identities here
+for reference.  The elementary commutation relations between spin
 operators are, with $\bmu\equiv-\mu\in\set{+1,-1}$ for brevity,
 \begin{align}
-  \f12 \sp{\sigma_j^z,\sigma_k^\mu}_-
-  &= \delta_{jk} \mu \sigma_j^\mu,
+  \f12 \sp{\sigma_\z^{(j)},\sigma_\mu^{(k)}}_-
+  &= \delta_{jk} \mu \sigma_\mu^{(j)},
   &
-  \sp{S_z,\sigma_j^\mu}_-
-  &= \f12 \sp{\sigma_j^z,S_\mu}_- = \mu \sigma_j^\mu,
+  \sp{S_\z,\sigma_\mu^{(j)}}_-
+  &= \f12 \sp{\sigma_\z^{(j)},S_\mu}_- = \mu \sigma_\mu^{(j)},
   &
-  \sp{S_z,S_\mu}_-
+  \sp{S_\z,S_\mu}_-
   &= \mu S_\mu,
   \label{eq:comm_z_base} \\
-  \sp{\sigma_j^\mu,\sigma_k^\bmu}_-
-  &= \delta_{jk} \mu \sigma_j^z,
+  \sp{\sigma_\mu^{(j)},\sigma_\bmu^{(k)}}_-
+  &= \delta_{jk} \mu \sigma_\z^{(j)},
   &
-  \sp{S_\mu,\sigma_j^\bmu}_-
-  &= \sp{\sigma_j^\mu,S_\bmu}_- = \mu \sigma_j^z,
+  \sp{S_\mu,\sigma_\bmu^{(j)}}_-
+  &= \sp{\sigma_\mu^{(j)},S_\bmu}_- = \mu \sigma_\z^{(j)},
   &
   \sp{S_\mu,S_\bmu}_-
-  &= 2 \mu S_z.
+  &= 2 \mu S_\z.
   \label{eq:comm_mu_base}
@@
 These relations can be used to inductively compute identities
-involving powers of collective spin operators.  Working through the
-possibilities, we find
-\begin{align}
-  S_z^m \sigma_j^\mu
-  = S_z^{m-1} \sigma_j^\mu \p{\mu + S_z}
-  = S_z^{m-2} \sigma_j^\mu \p{\mu + S_z}^2
+involving powers of collective spin operators.  By pushing through one
+spin operator at a time, we can find
+\begin{align}
+  S_\z^m \sigma_\mu^{(j)}
+  = S_\z^{m-1} \sigma_\mu^{(j)} \p{\mu + S_\z}
+  = S_\z^{m-2} \sigma_\mu^{(j)} \p{\mu + S_\z}^2
   = \cdots
-  = \sigma_j^\mu \p{\mu + S_z}^m,
+  = \sigma_\mu^{(j)} \p{\mu + S_\z}^m,
   \label{eq:push_z_mu_Ss}
@@
 \begin{align}
-  \sigma_j^z S_\mu^m
-  = \p{2\mu\sigma_j^\mu + S_\mu \sigma_j^z} S_\mu^{m-1}
-  = 2\mu\sigma_j^\mu S_\mu^{m-1} + S_\mu \sigma_j^z S_\mu^{m-1}
+  \sigma_\z^{(j)} S_\mu^m
+  = \p{2\mu\sigma_\mu^{(j)} + S_\mu \sigma_\z^{(j)}} S_\mu^{m-1}
+  = 2\mu\sigma_\mu^{(j)} S_\mu^{m-1} + S_\mu \sigma_\z^{(j)} S_\mu^{m-1}
   = \cdots
-  = 2\mu m\sigma_j^\mu S_\mu^{m-1} + S_\mu^m \sigma_j^z.
+  = 2\mu m\sigma_\mu^{(j)} S_\mu^{m-1} + S_\mu^m \sigma_\z^{(j)}.
   \label{eq:push_z_mu_sS}
@@
 \begin{align}
-  S_z^m S_\mu = S_\mu \p{\mu + S_z}^m,
-  &&
-  S_z S_\mu^m = S_\mu^m \p{\mu m + S_z},
+  S_\z^m S_\mu = S_\mu \p{\mu + S_\z}^m,
+  &&
+  S_\z S_\mu^m = S_\mu^m \p{\mu m + S_\z},
   \label{eq:push_z_mu_single}
 \end{align}
-where we can repeat the process of pushing through individual
+where we can repeat the process of pushing through individual $S_\z$
 operators $\ell$ times to get
 \begin{align}
-  S_z^\ell S_\mu^m
-  = S_z^{\ell-1} S_\mu^m \p{\mu m + S_z}
-  = S_z^{\ell-2} S_\mu^m \p{\mu m + S_z}^2
+  S_\z^\ell S_\mu^m
+  = S_\z^{\ell-1} S_\mu^m \p{\mu m + S_\z}
+  = S_\z^{\ell-2} S_\mu^m \p{\mu m + S_\z}^2
   = \cdots
-  = S_\mu^m \p{\mu m + S_z}^\ell.
+  = S_\mu^m \p{\mu m + S_\z}^\ell.
   \label{eq:push_z_mu}
@@
 
-\section{Commutator between powers of two transverse spin operators}
+\section{Commutation relations between powers of transverse spin
+  operators}
 \label{sec:comm_transverse}
@@
 \begin{align}
-  S_\mu^m \sigma_j^\bmu
-  &= S_\mu^{m-1}\sigma_j^\bmu S_\mu + \mu S_\mu^{m-1} \sigma_j^z \\
-  &= S_\mu^{m-2} \sigma_j^\bmu S_\mu^2 + \mu S_\mu^{m-2} \sigma_j^z S_\mu
-  + \mu S_\mu^{m-1} \sigma_j^z \\
-  &= \sigma_j^\bmu S_\mu^m
-  + \mu \sum_{k=0}^{m-1} S_\mu^k \sigma_j^z S_\mu^{m-k-1}
+  S_\mu^m \sigma_\bmu^{(j)}
+  &= S_\mu^{m-1}\sigma_\bmu^{(j)} S_\mu
+  + \mu S_\mu^{m-1} \sigma_\z^{(j)} \\
+  &= S_\mu^{m-2} \sigma_\bmu^{(j)} S_\mu^2
+  + \mu S_\mu^{m-2} \sigma_\z^{(j)} S_\mu
+  + \mu S_\mu^{m-1} \sigma_\z^{(j)} \\
+  &= \sigma_\bmu^{(j)} S_\mu^m
+  + \mu \sum_{k=0}^{m-1} S_\mu^k \sigma_\z^{(j)} S_\mu^{m-k-1}
   \label{eq:push_mu_Ss_start}.
@@
 While \eqref{eq:push_mu_Ss_start} gives us the commutator
-$\sp{S_\mu^m,\sigma_j^\bmu}_-$, we would like to enforce an ordering
-on products of spin operators, which will allow us to only keep track
-of operators that are linearly independent.  We choose to impose an
-ordering with all $\sigma_j^\bmu$ operators on the left, and all
-$\sigma_j^z$ operators on the right, as such an ordering will prove
-convenient for the calculations in this section.  This choice of
-ordering compels us to expand
-\begin{align}
-  \mu \sum_{k=0}^{m-1} S_\mu^k \sigma_j^z S_\mu^{m-k-1}
+$\sp{S_\mu^m,\sigma_\bmu^{(j)}}_-$, we would like to enforce an
+ordering on products of spin operators, which will allow us to only
+keep track of operators that are linearly independent.  We choose to
+impose an ordering with all $\sigma_\bmu^{(j)}$ operators on the left,
+and all $\sigma_\z^{(j)}$ operators on the right, as such an ordering
+will prove convenient for the calculations in this section.  This
+choice of ordering compels us to expand
+\begin{align}
+  \mu \sum_{k=0}^{m-1} S_\mu^k \sigma_\z^{(j)} S_\mu^{m-k-1}
   &= \mu \sum_{k=0}^{mq-1} S_\mu^k
-  \p{2\mu\sp{m-k-1} \sigma_j^\mu S_\mu^{m-k-2}
-    + S_\mu^{m-k-1} \sigma_j^z} \\
-  &= m \p{m-1} \sigma_j^\mu S_\mu^{m-2} + \mu m S_\mu^{m-1} \sigma_j^z,
+  \p{2\mu\sp{m-k-1} \sigma_\mu^{(j)} S_\mu^{m-k-2}
+    + S_\mu^{m-k-1} \sigma_\z^{(j)}} \\
+  &= m \p{m-1} \sigma_\mu^{(j)} S_\mu^{m-2}
+  + \mu m S_\mu^{m-1} \sigma_\z^{(j)},
 \end{align}
@@
 \begin{align}
-  S_\mu^m \sigma_j^\bmu
-  = \sigma_j^\bmu S_\mu^m + m \p{m-1} \sigma_j^\mu S_\mu^{m-2}
-  + \mu m S_\mu^{m-1} \sigma_j^z,
+  S_\mu^m \sigma_\bmu^{(j)}
+  = \sigma_\bmu^{(j)} S_\mu^m + m \p{m-1} \sigma_\mu^{(j)} S_\mu^{m-2}
+  + \mu m S_\mu^{m-1} \sigma_\z^{(j)},
   \label{eq:push_mu_Ss}
@@
   S_\mu^m S_\bmu = S_\bmu S_\mu^m
-  + m S_\mu^{m-1} \p{m - 1 + 2\mu S_z}.
+  + m S_\mu^{m-1} \p{m - 1 + 2\mu S_\z}.
   \label{eq:push_mu_single}
@@
   &= S_\mu^{m-1} S_\bmu^n S_\mu
-  + n \sp{S_\mu^{m-1} S_\bmu^{n-1} \p{1 - n + 2\mu S_z}} \\
+  + n \sp{S_\mu^{m-1} S_\bmu^{n-1} \p{1 - n + 2\mu S_\z}} \\
   &= S_\bmu^n S_\mu^m
   + n \sum_{k=0}^{m-1} S_\mu^{m-k-1} S_\bmu^{n-1}
-  \p{1 - n + 2\mu S_z} S_\mu^k,
+  \p{1 - n + 2\mu S_\z} S_\mu^k,
 \end{align}
@@
   \equiv n \sum_{k=0}^{m-1} S_\mu^{m-k-1} S_\bmu^{n-1}
-  \p{1 - n + 2\mu S_z} S_\mu^k.
+  \p{1 - n + 2\mu S_\z} S_\mu^k.
 \end{align}
 We now need rearrange the operators in $C_{mn;\mu}$ into a standard
-order, which means pushing all $S_z$ operators to the right and, for
+order, which means pushing all $S_\z$ operators to the right and, for
 the purposes of this calculation, all $S_\bmu$ operators on the left.
-We begin by pushing $S_\mu^k$ to the left of $S_z$, which takes
-$2\mu S_z\to 2\mu S_z+2k$, and then push $S_\mu^{m-k-1}$ to the right
-of $S_\bmu^{n-1}$, giving us
+We begin by pushing $S_\mu^k$ to the left of $S_\z$, which takes
+$2\mu S_\z\to 2\mu S_\z+2k$, and then push $S_\mu^{m-k-1}$ to the
+right of $S_\bmu^{n-1}$, giving us
 \begin{align}
@@
   \p{S_\bmu^{n-1} S_\mu^{m-k-1} + C_{m-k-1,n-1;\mu}} S_\mu^k
-  \p{2k + 1 - n + 2\mu S_z} \\
+  \p{2k + 1 - n + 2\mu S_\z} \\
   &= D_{mn;\mu}
   + n \sum_{k=0}^{m-2} C_{m-k-1,n-1;\mu}
-  S_\mu^k \p{2k + 1 - n + 2\mu S_z},
+  S_\mu^k \p{2k + 1 - n + 2\mu S_\z},
   \label{eq:C_mn}
 \end{align}
-where we have dropped the last term in the final sum because
-$C_{m-k-1,n-1;\mu}=0$ if $k=m-1$, and
+where we have dropped the last ($k=m-1$) term in the remaining sum
+because $C_{m-k-1,n-1;\mu}=0$ if $k=m-1$, and
 \begin{align}
   D_{mn;\mu}
-  \equiv mn S_\bmu^{n-1} S_\mu^{m-1} \p{m - n + 2\mu S_z}.
+  \equiv mn S_\bmu^{n-1} S_\mu^{m-1} \p{m - n + 2\mu S_\z}.
   \label{eq:D_mn}
 \end{align}
-To our despair, we have arrived in \eqref{eq:C_mn} to a recursive
-formula for $C_{mn;\mu}$.  Furthermore, we have not even managed to
-order the spin operators, as $C_{m-k-1,n-1;\mu}$ contains $S_z$
-operators which are to the left of $S_\mu^k$.  To sort all spin
+To our despair, we have arrived in \eqref{eq:C_mn} at a {\it
+  recursive} formula for $C_{mn;\mu}$.  Furthermore, we have not even
+managed to order all spin operators, as $C_{m-k-1,n-1;\mu}$ contains
+$S_\z$ operators which are to the left of $S_\mu^k$.  To sort all spin
 operators once and for all, we define
@@
   &= \p{m-k}n S_\bmu^{n-1} S_\mu^{m-k-1}
-  \p{m-k-n+2\mu S_z} S_\mu^k \\
-  &= \p{m-k}n S_\bmu^{n-1} S_\mu^{m-1} \p{k+m-n+2\mu S_z},
+  \p{m-k-n+2\mu S_\z} S_\mu^k \\
+  &= \p{m-k}n S_\bmu^{n-1} S_\mu^{m-1} \p{k+m-n+2\mu S_\z},
   \label{eq:D_mn_k}
@@
   &= D_{m-k,n;\mu} S_\mu^k + n \sum_{j=0}^{m-k-2}
-  C_{m-k-j-1,n-1;\mu} S_\mu^j \p{2j+1-n+2\mu S_z} S_\mu^k \\
+  C_{m-k-j-1,n-1;\mu} S_\mu^j \p{2j+1-n+2\mu S_\z} S_\mu^k \\
   &= D_{mn;\mu}^{(k)} + n \sum_{j=0}^{m-k-2}
-  C_{m-k-j-1,n-1;\mu} S_\mu^{j+k} \p{2j+2k+1-n+2\mu S_z} \\
+  C_{m-k-j-1,n-1;\mu} S_\mu^{j+k} \p{2j+2k+1-n+2\mu S_\z} \\
   &= D_{mn;\mu}^{(k)} + n \sum_{j=0}^{m-k-2}
-  C_{m-1,n-1;\mu}^{(k+j)} \p{2\sp{j+k}+1-n+2\mu S_z} \\
+  C_{m-1,n-1;\mu}^{(k+j)} \p{2\sp{j+k}+1-n+2\mu S_\z} \\
   &= D_{mn;\mu}^{(k)} + n \sum_{j=k}^{m-2}
-  C_{m-1,n-1;\mu}^{(j)} \p{2j+1-n+2\mu S_z}.
+  C_{m-1,n-1;\mu}^{(j)} \p{2j+1-n+2\mu S_\z}.
   \label{eq:C_mn_k}
@@
   E_{mn;\mu}^{(1)}
-  &= n \sum_{k=0}^{m-2} D_{m-1,n-1;\mu}^{(k)}
-  \p{2k+1-n+2\mu S_z}, \\
+  &= n \sum_{k=0}^{m-2} D_{m-1,n-1;\mu}^{(k)} \p{2k+1-n+2\mu S_\z}, \\
   E_{mn;\mu}^{(2)}
   &= n \sum_{k_1=0}^{m-2} \p{n-1} \sum_{k_2=k_1}^{m-3}
-  D_{m-2,n-2;\mu}^{(k_2)}
-  \p{2k_2+2-n+2\mu S_z} \p{2k_1+1-n+2\mu S_z},
+  D_{m-2,n-2;\mu}^{(k_2)} \p{2k_2+2-n+2\mu S_\z} \p{2k_1+1-n+2\mu S_\z},
 \end{align}
@@
   \sum_{k_1=0}^{m-2} \sum_{k_2=k_1}^{m-3} \cdots\sum_{k_p=k_{p-1}}^{m-p-1}
-  D_{m-p,n-p;\mu}^{(k_p)} \prod_{j=1}^p \p{2k_j+j-n+2\mu S_z}.
+  D_{m-p,n-p;\mu}^{(k_p)} \prod_{j=1}^p \p{2k_j+j-n+2\mu S_\z}.
   \label{eq:E_mn_p}
 \end{align}
-In principle, the expressions in \eqref{eq:D_mn} and
+In principle, the expressions in \eqref{eq:D_mn_k} and
 \eqref{eq:C_mn_E}-\eqref{eq:E_mn_p} can be used to evaluate the
@@
   \sum_{\ell=0}^{m-p-1} D_{m-p,n-p;\mu}^{(\ell)} \p{2\ell+F_{np;\mu}}
-  \sum_{\p{\v k,p,\ell}} \prod_{j=1}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}},
+  \sum_{\p{\v k,p-1,\ell}} \prod_{j=1}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}},
   \label{eq:E_mn_p_sum}
@@
 \begin{align}
-  F_{np;\mu} \equiv p - n + 2\mu S_z,
-  &&
-  \sum_{\p{\v k,p,\ell}} X \equiv
+  F_{np;\mu} \equiv p - n + 2\mu S_\z,
+  &&
+  \sum_{\p{\v k,q,\ell}} X \equiv
   \sum_{k_1=0}^\ell \sum_{k_2=k_1}^\ell
-  \cdots \sum_{k_{p-1}=k_{p-2}}^\ell X.
-\end{align}
-We now define
+  \cdots \sum_{k_q=k_{q-1}}^\ell X.
+\end{align}
+We now further define
 \begin{align}
@@
 \begin{align}
-  \sum_{\p{\v k,p,\ell}} \prod_{j=1}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}}
-  &= \sum_{\p{\v k,p-1,\ell}}
+  \sum_{\p{\v k,p-1,\ell}} \prod_{j=1}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}}
+  &= \sum_{\p{\v k,p-2,\ell}}
   \prod_{j=2}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}}
@@
   &= \f1{r!} \sum_{\p{\v k,p-r,\ell}}
-  \prod_{j=1+r}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}}
-  \prod_{q=1}^r f_{np\ell;\mu}\p{k_{p-1-r},q} \\
+  \prod_{j=r}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}}
+  \prod_{q=1}^{r-1} f_{np\ell;\mu}\p{k_{p-r},q} \\
   &= \f1{\p{p-1}!} \prod_{q=1}^{p-1} f_{np\ell;\mu}\p{0,q} \\
@@
 Substitution of this result together with $D_{m-p,n-p;\mu}^{(\ell)}$
-with \eqref{eq:D_mn_k} into \eqref{eq:E_mn_p_sum} then gives us
+using \eqref{eq:D_mn_k} into \eqref{eq:E_mn_p_sum} then gives us
 \begin{align}
@@
 \end{align}
-We can further expand
+We can further simplify
 \begin{align}
   \prod_{q=0}^p \p{m-p-q+F_{np;\mu}}
-  = \prod_{q=0}^p \p{m-n-q+2\mu S_z}
-  = \sum_{q=0}^{p+1} \p{-1}^{p+1-q} { p+1 \brack q } \p{m-n+2\mu S_z}^q,
+  = \prod_{q=0}^p \p{m-n-q+2\mu S_\z}
+  = \sum_{q=0}^{p+1} \p{-1}^{p+1-q}
+  { p+1 \brack q } \p{m-n+2\mu S_\z}^q,
 \end{align}
 where ${ n \brack k }$ is an unsigned Stirling number of the first
-kind, and finally
-\begin{align}
-  \sum_{q=0}^p \p{-1}^{p-q} { p \brack q } \p{m-n+2\mu S_z}^q
+kind, i.e.~the number of ways to pick a set of $k$ elements from a set
+of $n$ elements, and finally
+\begin{align}
+  \sum_{q=0}^p \p{-1}^{p-q} { p \brack q } \p{m-n+2\mu S_\z}^q
   &= \sum_{q=0}^p \p{-1}^{p-q} { p \brack q } \sum_{\ell=0}^q
-  { q \choose \ell } \p{m-n}^{q-\ell} \p{2\mu S_z}^\ell \\
+  { q \choose \ell } \p{m-n}^{q-\ell} \p{2\mu S_\z}^\ell \\
   &= \sum_{\ell=0}^p \p{2\mu}^\ell \sum_{q=\ell}^p \p{-1}^{p-q}
-  { p \brack q } { q \choose \ell } \p{m-n}^{q-\ell} S_z^\ell.
+  { p \brack q } { q \choose \ell } \p{m-n}^{q-\ell} S_\z^\ell.
 \end{align}
@@
   S_\bmu^{n-p} S_\mu^{m-p} \sum_{\ell=0}^p
-  \mu^\ell \epsilon_{mnp}^{(\ell)} S_z^\ell,
+  \mu^\ell \epsilon_{mnp}^{(\ell)} S_\z^\ell,
 \end{align}
@@
   p! { m \choose p } { n \choose p } S_\bmu^{n-p} S_\mu^{m-p}
-  \sum_{\ell=0}^p \mu^\ell \epsilon_{mnp}^{(\ell)} S_z^\ell,
+  \sum_{\ell=0}^p \mu^\ell \epsilon_{mnp}^{(\ell)} S_\z^\ell,
   \label{eq:comm_mu}
@@
   p! { m \choose p } { n \choose p } S_\bmu^{n-p} S_\mu^{m-p}
-  \sum_{\ell=0}^p \mu^\ell \epsilon_{mnp}^{(\ell)} S_z^\ell,
-  \label{eq:push_mu_nu}
-\end{align}
-If we wish to order collective spin operators with $S_z$ in between
-$S_\bmu$ and $S_\mu$, then we find
+  \sum_{\ell=0}^p \mu^\ell \epsilon_{mnp}^{(\ell)} S_\z^\ell,
+  \label{eq:push_mu_bmu}
+\end{align}
+If we wish to order products of collective spin operators with $S_\z$
+in between $S_\bmu$ and $S_\mu$, then
 \begin{align}
@@
   \equiv \sum_{\ell=0}^p \mu^\ell \epsilon_{mnp}^{(\ell)}
-  \p{-\mu\sp{m-p} + S_z}^\ell
-  = \sum_{q=0}^p \mu^q \xi_{mnp}^{(q)} S_z^q,
+  \p{-\mu\sp{m-p} + S_\z}^\ell
+  = \sum_{q=0}^p \mu^q \xi_{mnp}^{(q)} S_\z^q,
   \label{eq:Z_mn}
@@
 
-\section{Commutator between general ordered collective spin operators}
+\section{Commutation relations between ordered products of collective
+  spin operators}
 \label{sec:comm_general}
@@
 If all collective spin operators are written in the form
-$S_\mu^\ell S_z^m S_\bmu^n$ for $\bmu\equiv-\mu$, then the most
-general commutator we need to consider is
-\begin{align}
-  \sp{S_\mu^p S_z^a S_\bmu^q, S_\mu^r S_z^b S_\bmu^s}_-
-  = \S^{paq}_{rbs;\mu} - \S^{rbs}_{paq;\mu},
+$S_\mu^\ell S_\z^m S_\bmu^n$ for $\bmu\equiv-\mu\in\set{+1,-1}$, then
+the most general (anti-)commutator we need to consider is
+\begin{align}
+  \sp{S_\mu^p S_\z^a S_\bmu^q, S_\mu^r S_\z^b S_\bmu^s}_\pm
+  = \S^{paq}_{rbs;\mu} \pm \S^{rbs}_{paq;\mu},
 \end{align}
@@
   \S^{paq}_{rbs;\mu}
-  \equiv S_\mu^p S_z^a S_\bmu^q S_\mu^r S_z^b S_\bmu^s
+  \equiv S_\mu^p S_\z^a S_\bmu^q S_\mu^r S_\z^b S_\bmu^s
   = \sum_{k=0}^{\min\set{q,r}} k! { q \choose k } { r \choose k }
-  S_\mu^{p+r-k} \tilde Z_{qrab;\bmu}^{(k)} S_\bmu^{q+s-k},
+  S_\mu^{p+r-k} \tilde Z_{qrab;\mu}^{(k)} S_\bmu^{q+s-k},
 \end{align}
@@
   \tilde Z_{qrab;\mu}^{(k)}
-  &\equiv \p{\mu\sp{r-k}+S_z}^a
-  Z_{qr;\bmu}^{(k)} \p{\mu\sp{q-k}+S_z}^b \\
+  &\equiv \p{\mu\sp{r-k}+S_\z}^a
+  Z_{qr;\bmu}^{(k)} \p{\mu\sp{q-k}+S_\z}^b \\
   &= \sum_{\ell=0}^k \p{-1}^\ell \xi_{qrk}^{(\ell)}
@@
   \sum_{n=0}^b \p{q-k}^{b-n} { b \choose n }
-  \mu^{a+b+\ell-m-n} S_z^{\ell+m+n}.
-\end{align}
-
-\section{Decoherence of collective spin operators from uncorrelated
-  spin decay}
-\label{sec:decoherence}
-
-The decoherence operator for uncorrelated decay ($\mu=-1$) or
-excitation ($\mu=1$) of individual spins is, with $\bmu\equiv-\mu$,
-\begin{align}
-  \D_\mu\p{\O}
-  \equiv \sum_j\p{\sigma_j^\bmu \O \sigma_j^\mu
-    - \f12\sp{\sigma_j^\bmu \sigma_j^\mu,\O}_+},
+  \mu^{a+b+\ell-m-n} S_\z^{\ell+m+n}.
+\end{align}
+
+\section{Sandwich identities for single-spin decoherence calculations}
+\label{sec:sandwich_single}
+
+In this section we derive several identities which will be necessary
+for computing the effects of single-spin decoherence on ordered
+products of collective spin operators.  These identities all involve
+sandwiching a collective spin operator between operators that act on
+individual spins only, and summing such terms over all individual spin
+indices.  Our general strategy will be to use commutation relations to
+push single-spin operators together, and then evaluate the sum to
+arrive at an expression in terms of collective spin operators only.
+
+We first compute sums of single-spin operators sandwiching $S_\z^m$,
+which will in turn be used to compute similar sums sandwiching
+$S_\mu^\ell S_\z^m S_\bmu^n$.  The simplest case is
+\begin{align}
+  \sum_j \sigma_\z^{(j)} S_\z^m \sigma_\z^{(j)}
+  = \sum_j \sigma_\z^{(j)} \sigma_\z^{(j)} S_\z^m
+  = \sum_j \1_j S_\z^m
+  = 2 S S_\z^m.
+\end{align}
+Similarly,
+\begin{align}
+  \sum_j \sigma_\z^{(j)} S_\z^m \sigma_\mu^{(j)}
+  = S_\z^m \sum_j \sigma_\z^{(j)} \sigma_\mu^{(j)}
+  = S_\z^m \sum_j \mu \sigma_\mu^{(j)}
+  = \mu S_\z^m S_\mu
+  = \mu S_\mu \p{\mu+S_\z}^m,
+\end{align}
+where in the last equality we used the identity derived in
+\eqref{eq:push_z_mu_single} to push $S_\mu$ through $S_\z^m$.  Using
+identity \eqref{eq:push_z_mu_Ss} to push $\sigma_\mu^{(j)}$ through
+$S_\z^m$, we can also find
+\begin{align}
+  \sum_j \sigma_\mu^{(j)} S_\z^m \sigma_\mu^{(j)}
+  = \sum_j \sigma_\mu^{(j)} \sigma_\mu^{(j)} \p{\mu+S_\z}^m
+  = 0,
+\end{align}
+and
+\begin{align}
+  \sum_j \sigma_\bmu^{(j)} S_\z^m \sigma_\mu^{(j)}
+  = \sum_j \sigma_\bmu^{(j)} \sigma_\mu^{(j)} \p{\mu+S_\z}^m
+  = \p{S-\mu S_\z} \p{\mu+S_\z}^m.
+\end{align}
+We are now equipped to derive similar identities for general ordered
+products of collective spin operators.  For quick reference, we
+provide here identity \eqref{eq:push_z_mu_sS} and its conjugate:
+\begin{align}
+  \sigma_\z^{(j)} S_\mu^\ell
+  = S_\mu^\ell \sigma_\z^{(j)} + 2\mu \ell S_\mu^{\ell-1}\sigma_\mu^{(j)},
+  &&
+  S_\bmu^n \sigma_\z^{(j)}
+  = \sigma_\z^{(j)} S_\bmu^n + 2\mu n \sigma_\bmu^{(j)} S_\bmu^{n-1},
+\end{align}
+and likewise with identity \eqref{eq:push_mu_Ss}:
+\begin{align}
+  \sigma_\bmu^{(j)} S_\mu^\ell
+  &= S_\mu^\ell \sigma_\bmu^{(j)}
+  - \ell \p{\ell-1} S_\mu^{\ell-2} \sigma_\mu^{(j)}
+  - \mu \ell S_\mu^{\ell-1} \sigma_\z^{(j)}, \\
+  S_\bmu^n \sigma_\mu^{(j)}
+  &= \sigma_\mu^{(j)} S_\bmu^n
+  - n \p{n-1} \sigma_\bmu^{(j)} S_\bmu^{n-2}
+  - \mu n \sigma_\z^{(j)} S_\bmu^{n-1}.
+\end{align}
+Working through the combinations, we then compute
+\begin{align}
+  \sum_j \sigma_\mu^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\bmu^{(j)}
+  &= S_\mu^\ell \p{S+\mu S_\z}\p{-\mu+S_\z}^m S_\bmu^n,
+  \label{eq:sand_mu_bmu} \allowdisplaybreaks \\
+  \sum_j \sigma_\mu^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\z^{(j)}
+  &= -\mu S_\mu^{\ell+1} S_\z^m S_\bmu^n
+  + 2\mu n S_\mu^\ell \p{S+\mu S_\z} \p{-\mu+S_\z}^m S_\bmu^{n-1},
+  \allowdisplaybreaks \\
+  \sum_j \sigma_\z^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\bmu^{(j)}
+  &= -\mu S_\mu^\ell S_\z^m S_\bmu^{n+1}
+  + 2\mu \ell S_\mu^{\ell-1} \p{S+\mu S_\z}\p{-\mu+S_\z}^m S_\bmu^n,
+  \allowdisplaybreaks \\
+  \sum_j \sigma_\z^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\z^{(j)}
+  &= 2\p{S-\ell-n} S_\mu^\ell S_\z^m S_\bmu^n
+  + 4\ell n S_\mu^{\ell-1} \p{S+\mu S_\z}
+  \p{-\mu+S_\z}^m S_\bmu^{n-1},
+  \label{eq:sand_z_z} \allowdisplaybreaks \\
+  \sum_j \sigma_\mu^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\mu^{(j)}
+  &= -n\p{n-1} S_\mu^\ell \p{S+\mu S_\z} \p{-\mu+S_\z}^m S_\bmu^{n-2}
+  + n S_\mu^{\ell+1} S_\z^m S_\bmu^{n-1},
+  \allowdisplaybreaks \\
+  \sum_j \sigma_\bmu^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\bmu^{(j)}
+  &= -\ell \p{\ell-1} S_\mu^{\ell-2}
+  \p{S+\mu S_\z} \p{-\mu+S_\z}^m S_\bmu^n
+  + \ell S_\mu^{\ell-1} S_\z^m S_\bmu^{n+1},
+  \allowdisplaybreaks \\
+  \sum_j \sigma_\z^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\mu^{(j)}
+  &= \mu S_\mu^{\ell+1} \p{\mu+S_\z}^m S_\bmu^n
+  + \mu n \p{-2S+2\ell+n-1} S_\mu^\ell S_\z^m S_\bmu^{n-1} \notag \\
+  &\qquad - 2\mu\ell n\p{n-1} S_\mu^{\ell-1}
+  \p{S+\mu S_\z} \p{-\mu+S_\z}^m S_\bmu^{n-2},
+  \allowdisplaybreaks \\
+  \sum_j \sigma_\bmu^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\z^{(j)}
+  &= \mu S_\mu^\ell \p{\mu+S_\z}^m S_\bmu^{n+1}
+  + \mu \ell \p{-2S+2n+\ell-1} S_\mu^{\ell-1} S_\z^m S_\bmu^n \notag \\
+  &\qquad - 2\mu\ell n\p{\ell-1} S_\mu^{\ell-2}
+  \p{S+\mu S_\z} \p{-\mu+S_\z}^m S_\bmu^{n-1},
+  \allowdisplaybreaks \\
+  \sum_j \sigma_\bmu^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\mu^{(j)}
+  &= S_\mu^\ell \p{S - \ell - n - \mu S_\z} \p{\mu + S_\z}^m S_\bmu^n
+  + \ell n \p{2S - \ell - n + 2}
+  S_\mu^{\ell-1} S_\z^m S_\bmu^{n-1} \notag \\
+  &\qquad + \ell n \p{\ell-1} \p{n-1} S_\mu^{\ell-2} \p{S + \mu S_\z}
+  \p{-\mu + S_\z}^m S_\bmu^{n-2}.
+  \label{eq:sand_bmu_mu}
+\end{align}
+
+\section{Uncorrelated, permutationally-symmetric single-spin
+  decoherence}
+\label{sec:decoherence_single}
+
+In this section we work out the effects of permutationally-symmetric
+decoherence of individual spins on ordered products of collective spin
+operators.  For shorthand, we define
+\begin{align}
+  \D\p{\gamma;\O} \equiv \D\p{\set{\gamma^{(j)}};\O}
+  = \sum_j\p{{\gamma^{(j)}}^\dag \O \gamma^{(j)}
+    - \f12\sp{{\gamma^{(j)}}^\dag \gamma^{(j)}, \O}_+},
+\end{align}
+where $\gamma$ is an operator that acts on a single spin, and
+$\gamma^{(j)}$ is an operator that acts with $\gamma$ on spin $j$ and
+the identity on all other spins.
+
+\subsection{Decay-type decoherence}
+\label{sec:decay_single}
+
+The effect of decoherence via uncorrelated decay ($\mu=-1$) or
+excitation ($\mu=1$) of individual spins is described by (with
+$\bmu\equiv-\mu$),
+\begin{align}
+  \D\p{\sigma_\mu; \O}
+  = \sum_j\p{\sigma_\bmu^{(j)} \O \sigma_\mu^{(j)}
+    - \f12\sp{\sigma_\bmu^{(j)} \sigma_\mu^{(j)},\O}_+}
+  = \sum_j \sigma_\bmu^{(j)} \O \sigma_\mu^{(j)}
+    - S \O + \f{\mu}{2} \sp{S_\z, \O}_+.
+\end{align}
+In order to determine the effect of this decoherence on ordered
+products of collective spin operators, we expand the anti-commutator
+\begin{align}
+  \sp{S_\z, S_\mu^\ell S_\z^m S_\bmu^n}_+
+  = S_\z S_\mu^\ell S_\z^m S_\bmu^n + S_\mu^\ell S_\z^m S_\bmu^n S_\z
+  = S_\mu^\ell\p{\mu\ell + \mu n + 2 S_\z} S_\z^m S_\bmu^n,
+\end{align}
+which implies, using \eqref{eq:sand_mu_bmu},
+\begin{align}
+  \D\p{\sigma_\bmu; S_\mu^\ell S_\z^m S_\bmu^n}
+  &= S_\mu^\ell \p{S+\mu S_\z}\p{-\mu+S_\z}^m S_\bmu^n
+  - S_\mu^\ell\p{S + \f12\sp{\ell+n} + \mu S_\z} S_\z^m S_\bmu^n \\
+  &= S_\mu^\ell \sp{\p{S+\mu S_\z}
+    \sum_{k=0}^{m-1} \p{-\mu}^{m-k} { m \choose k } S_\z^k
+    - \f12 \p{\ell+n} S_\z^m} S_\bmu^n,
+  \label{eq:decay_diff}
+\end{align}
+and, using \eqref{eq:sand_bmu_mu},
+\begin{align}
+  \D\p{\sigma_\mu; S_\mu^\ell S_\z^m S_\bmu^n}
+  &= S_\mu^\ell \p{S - \ell - n - \mu S_\z} \p{\mu + S_\z}^m S_\bmu^n
+  + \ell n \p{2S - \ell - n + 2}
+  S_\mu^{\ell-1} S_\z^m S_\bmu^{n-1} \notag \\
+  &\qquad + \ell n \p{\ell-1} \p{n-1} S_\mu^{\ell-2} \p{S + \mu S_\z}
+  \p{-\mu + S_\z}^m S_\bmu^{n-2} \notag \\
+  &\qquad - S_\mu^\ell\p{S - \f12\sp{\ell+n} - \mu S_\z}
+  S_\z^m S_\bmu^n \allowdisplaybreaks \\
+  &= S_\mu^\ell \sp{\p{S-\ell-n-\mu S_\z}
+    \sum_{k=0}^{m-1} \mu^{m-k} { m \choose k } S_\z^k
+    - \f12 \p{\ell+n} S_\z^m} S_\bmu^n \notag \\
+  &\qquad + \ell n \p{2S - \ell - n + 2}
+  S_\mu^{\ell-1} S_\z^m S_\bmu^{n-1} \notag \\
+  &\qquad + \ell n \p{\ell-1} \p{n-1} S_\mu^{\ell-2} \p{S + \mu S_\z}
+  \sum_{k=0}^m \p{-\mu}^{m-k} { m \choose k } S_\z^k S_\bmu^{n-2}.
+  \label{eq:decay_same}
+\end{align}
+Decoherence via jump operators $\sigma_\bmu^{(j)}$ only couples
+operators $S_\mu^\ell S_\z^m S_\bmu^n$ to operators
+$S_\mu^\ell S_\z^{m'} S_\bmu^n$ with $m'\le m$.  Decoherence via jump
+operators $\sigma_\mu^{(j)}$, meanwhile, makes operators
+$S_\mu^\ell S_\z^m S_\bmu^n$ ``grow'' in $S_\z$ through the last term
+in \eqref{eq:decay_same} with $k=m$, although the sum $\ell+m+n$ does
+not grow.
+
+\subsection{Dephasing}
+\label{sec:dephasing_single}
+
+The effect of decoherence via single-spin dephasing is described by
+\begin{align}
+  \D\p{\sigma_\z/\sqrt2; \O}
+  = \sum_j\p{\sigma_\z^{(j)} \O \sigma_\z^{(j)}
+    - \f12\sp{\sigma_\z^{(j)} \sigma_\z^{(j)},\O}_+}
+  = \f12 \sum_j \sigma_\z^{(j)} \O \sigma_\z^{(j)} - S \O,
+\end{align}
+where the factor $\sqrt2$ is inserted as
+e.g.~$\D\p{\sigma_\z/\sqrt2;S_\x}=-S_\x$.  From \eqref{eq:sand_z_z},
+we then have
+\begin{align}
+  \D\p{\sigma_\z/\sqrt2; S_\mu^\ell S_\z^m S_\bmu^n}
+  &= -\p{\ell+n} S_\mu^\ell S_\z^m S_\bmu^n
+  + 2 \ell n S_\mu^{\ell-1} \p{S + \mu S_\z}
+  \p{-\mu + S_\z}^m S_\bmu^{n-1} \\
+  &= -\p{\ell+n} S_\mu^\ell S_\z^m S_\bmu^n
+  + 2 \ell n S_\mu^{\ell-1} \p{S + \mu S_\z}
+  \sum_{k=0}^m \p{-\mu}^{m-k} { m \choose k } S_\z^k S_\bmu^{n-1}.
+\end{align}
+Decoherence via single-spin dephasing also makes operators
+$S_\mu^\ell S_\z^m S_\bmu^n$ ``grow'' in $S_\z$, although the sum
+$\ell+m+n$ still does not grow.
+
+\subsection{The general case}
+\label{sec:general_single}
+
+The most general type of single-spin decoherence is described by
+\begin{align}
+  \D\p{\gamma;\O}
+  = \sum_j\p{{\gamma^{(j)}}^\dag \O \gamma^{(j)}
+    - \f12\sp{{\gamma^{(j)}}^\dag \gamma^{(j)}, \O}_+},
+  &&
+  \gamma \equiv \gamma_\z \sigma_\z
+  + \gamma_+ \sigma_+ + \gamma_- \sigma_-.
+  \label{eq:D_general_single_start}
+\end{align}
+To simplify \eqref{eq:D_general_single_start}, we expand
+\begin{align}
+  \gamma^\dag \O \gamma
+  = \abs{\gamma_\z}^2 \sigma_\z \O \sigma_\z
+  + \sum_\mu \p{\abs{\gamma_\mu}^2 \sigma_\bmu \O \sigma_\mu
+    + \gamma_\bmu^* \gamma_\mu \sigma_\mu \O \sigma_\mu
+    + \gamma_\z^* \gamma_\mu \sigma_\z \O \sigma_\mu
+    + \gamma_\bmu^* \gamma_\z \sigma_\mu \O \sigma_\z},
+\end{align}
+and
+\begin{align}
+  \gamma^\dag \gamma
+  = \abs{\gamma_\z}^2
+  + \sum_\mu \p{\abs{\gamma_\mu}^2 \f12 \sp{1-\mu\sigma_\z}
+    + \mu \sp{\gamma_\z^*\gamma_\mu - \gamma_\bmu^*\gamma_\z}
+    \sigma_\mu},
+\end{align}
+which implies
+\begin{align}
+  \D\p{\gamma;\O}
+  &= \sum_{X\in\set{\z,+,-}} \abs{\gamma_X}^2 \D\p{\sigma_X;\O}
+  + \sum_{j,\mu}
+  \p{\gamma_\bmu^* \gamma_\mu \sigma_\mu^{(j)} \O \sigma_\mu^{(j)}
+    + \gamma_\z^* \gamma_\mu \sigma_\z^{(j)} \O \sigma_\mu^{(j)}
+    + \gamma_\bmu^* \gamma_\z \sigma_\mu^{(j)} \O \sigma_\z^{(j)}}
+  \notag \\
+  &\qquad -\f12 \sum_\mu \mu
+  \p{\gamma_\z^*\gamma_\mu - \gamma_\bmu^*\gamma_\z} \sp{S_\mu, \O}_+.
+  \label{eq:D_general_single}
+\end{align}
+In order to compute the effect of this decoherence on ordered products
+of collective spin operators, we expand the anti-commutators
+\begin{align}
+  \sp{S_\mu, S_\mu^\ell S_\z^m S_\bmu^n}_+
+  &= S_\mu^{\ell+1} \sp{S_\z^m+\p{\mu+S_\z}^m} S_\bmu^n
+  - n S_\mu^\ell \p{n-1+2\mu S_\z} S_\z^m S_\bmu^{n-1}, \\
+  \sp{S_\bmu, S_\mu^\ell S_\z^m S_\bmu^n}_+
+  &= S_\mu^\ell \sp{S_\z^m+\p{\mu+S_\z}^m} S_\bmu^{n+1}
+  - \ell S_\mu^{\ell-1} \p{\ell-1+2\mu S_\z} S_\z^m S_\bmu^n.
+  \label{eq:S_mu_acomm}
+\end{align}
+These anti-commutators together with the results in Sections
+\ref{sec:sandwich_single}, \ref{sec:decay_single}, and
+\ref{sec:dephasing_single} are sufficient to expand
+\eqref{eq:D_general_single} in terms of ordered products of collective
+spin operators.
+
+\section{Sandwich identities for collective-spin decoherence
+  calculations}
+\label{sec:sandwich_collective}
+
+In analogy with the work in Appendix \ref{sec:sandwich_single}, in
+this section we work out sandwich identities necessary for
+collective-spin decoherence calculations.  The simplest cases are
+\begin{align}
+  S_\mu S_\mu^\ell S_\z^m S_\bmu^n S_\bmu
+  &= S_\mu^{\ell+1} S_\z^m S_\bmu^{n+1},
+  \allowdisplaybreaks \\
+  S_\mu S_\mu^\ell S_\z^m S_\bmu^n S_\z
+  &= S_\mu^{\ell+1} \p{\mu n+S_\z} S_\z^m S_\bmu^n,
+  \allowdisplaybreaks \\
+  S_\z S_\mu^\ell S_\z^m S_\bmu^n S_\bmu
+  &= S_\mu^\ell \p{\mu\ell+S_\z} S_\z^m S_\bmu^{n+1},
+  \allowdisplaybreaks \\
+  S_\z S_\mu^\ell S_\z^m S_\bmu^n S_\z
+  &= S_\mu^\ell \p{\ell n + \mu\sp{\ell+n} S_\z + S_\z^2}
+  S_\z^m S_\bmu^n.
+\end{align}
+With a bit more work, we can also find
+\begin{align}
+  S_\mu^\ell S_\z^m S_\bmu^n S_\mu
+  &= S_\mu^{\ell+1} \p{\mu+S_\z}^m S_\bmu^n
+  - n S_\mu^\ell \p{n-1+2\mu S_\z} S_\z^m S_\bmu^{n-1},
+  \allowdisplaybreaks \\
+  S_\bmu S_\mu^\ell S_\z^m S_\bmu^n
+  &= S_\mu^\ell \p{\mu+S_\z}^m S_\bmu^{n+1}
+  - \ell S_\mu^{\ell-1} \p{\ell-1+2\mu S_\z} S_\z^m S_\bmu^n,
+\end{align}
+which implies
+\begin{align}
+  S_\mu S_\mu^\ell S_\z^m S_\bmu^n S_\mu
+  &= S_\mu^{\ell+2} \p{\mu+S_\z}^m S_\bmu^n
+  - n S_\mu^{\ell+1} \p{n-1+2\mu S_\z} S_\z^m S_\bmu^{n-1},
+  \allowdisplaybreaks \\
+  S_\bmu S_\mu^\ell S_\z^m S_\bmu^n S_\bmu
+  &= S_\mu^\ell \p{\mu+S_\z}^m S_\bmu^{n+2}
+  - \ell S_\mu^{\ell-1} \p{\ell-1+2\mu S_\z} S_\z^m S_\bmu^{n+1},
+  \allowdisplaybreaks \\
+  S_\z S_\mu^\ell S_\z^m S_\bmu^n S_\mu
+  &= S_\mu^{\ell+1} \p{\mu\sp{\ell+1}+S_\z} \p{\mu+S_\z}^m S_\bmu^n
+  \notag \\
+  &\qquad - n S_\mu^\ell
+  \p{\mu\ell\sp{n-1} + \sp{2\ell+n-1}S_\z + 2\mu S_\z^2}
+  S_\z^m S_\bmu^{n-1},
+  \allowdisplaybreaks \\
+  S_\bmu S_\mu^\ell S_\z^m S_\bmu^n S_\z
+  &= S_\mu^\ell \p{\mu\sp{n+1}+S_\z} \p{\mu+S_\z}^m S_\bmu^{n+1}
+  \notag \\
+  &\qquad - \ell S_\mu^{\ell-1}
+  \p{\mu n\sp{\ell-1} + \sp{2n+\ell-1}S_\z + 2\mu S_\z^2}
+  S_\z^m S_\bmu^n.
+\end{align}
+Finally, we compute
+\begin{align}
+  S_\bmu S_\mu^\ell S_\z^m S_\bmu^n S_\mu
+  &= \sp{S_\mu^\ell S_\bmu - \ell S_\mu^{\ell-1} \p{\ell-1+2\mu S_\z}}
+  S_\z^m
+  \sp{S_\mu S_\bmu^n - n \p{n-1+2\mu S_\z} S_\bmu^{n-1}} \\
+  &= S_\mu^\ell S_\bmu S_\z^m S_\mu S_\bmu^n
+  - S_\mu^\ell \p{\ell\sp{\ell+1} + n\sp{n+1}+2\mu\sp{\ell+n} S_\z}
+  \p{\mu+S_\z}^m S_\bmu^n \notag \\
+  &\qquad + \ell n S_\mu^{\ell-1}
+  \p{\sp{\ell-1}\sp{n-1}+2\mu\sp{\ell+n-2} S_\z + 4 S_\z^2}
+  S_\z^m S_\bmu^{n-1},
 \end{align}
@@
 \begin{align}
-  -\f12 \sp{\sigma_j^\bmu \sigma_j^\mu,\O}_+
-  = -\f12\sp{\p{S-\mu S_z}, \O}_+
-  = -S\O + \f{\mu}{2} \sp{S_z, \O}_+,
+  S_\bmu S_\z^m S_\mu
+  = S_\bmu S_\mu \p{\mu+S_\z}^m
+  = \p{S_\mu S_\bmu - 2\mu S_\z} \p{\mu+S_\z}^m
+  = S_\mu \p{2\mu + S_\z}^m S_\bmu - 2\mu S_\z \p{\mu + S_\z}^m,
 \end{align}
@@
 \begin{align}
-  \D_\mu\p{\O}
-  = \sum_j\sigma_j^\bmu \O \sigma_j^\mu
-    - S \O + \f{\mu}{2} \sp{S_z, \O}_+.
-  \label{eq:D_mu}
-\end{align}
-
-\subsection{The simple case}
-
-If we consider only symmetrically ordered collective spin operators,
-there are two cases to work out: the effect of decoherence through
-$\D_\mu$ on $S_\mu^\ell S_z^m S_\bmu^n$, and the effect on
-$S_\bmu^\ell S_z^m S_\mu^n$.  We work out the latter case first,
-namely
-\begin{align}
-  \D_\mu\p{S_\bmu^\ell S_z^m S_\mu^n}
-  = \sum_j\sigma_j^\bmu S_\bmu^\ell S_z^m S_\mu^n \sigma_j^\mu
-  - S S_\bmu^\ell S_z^m S_\mu^n
-  + \f{\mu}{2} \sp{S_z, S_\bmu^\ell S_z^m S_\mu^n}_+.
-\end{align}
-To simplify the first term, we use \eqref{eq:comm_z_base} to push the
-jump operators together, finding
-\begin{align}
-  \sum_j \sigma_j^\bmu S_\bmu^\ell S_z^m S_\mu^n \sigma_j^\mu
-  = S_\bmu^\ell \sum_j \sigma_j^\bmu \sigma_j^\mu \p{\mu+S_z}^m S_\mu^n
-  = S_\bmu^\ell \p{S+S_z} \p{\mu+S_z}^m S_\mu^n.
-\end{align}
-We can likewise push $S_z$ operators together using
-\eqref{eq:push_z_mu_single} to expand
-\begin{align}
-  \sp{S_z, S_\bmu^\ell S_z^m S_\mu^n}_+
-  = S_z S_\bmu^\ell S_z^m S_\mu^n + S_\bmu^\ell S_z^m S_\mu^n S_z
-  = S_\bmu^\ell\p{\bmu\ell - \mu n + 2 S_z} S_z^m S_\mu^n,
+  S_\bmu S_\mu^\ell S_\z^m S_\bmu^n S_\mu
+  &= S_\mu^{\ell+1} \p{2\mu + S_\z}^m S_\bmu^{n+1}
+  - S_\mu^\ell \p{\ell\sp{\ell+1} + n\sp{n+1}+2\mu\sp{\ell+n+1} S_\z}
+  \p{\mu+S_\z}^m S_\bmu^n \notag \\
+  &\qquad + \ell n S_\mu^{\ell-1}
+  \p{\sp{\ell-1}\sp{n-1}+2\mu\sp{\ell+n-2} S_\z + 4 S_\z^2}
+  S_\z^m S_\bmu^{n-1}.
+\end{align}
+
+\section{Collective spin decoherence}
+\label{sec:decoherence_collective}
+
+In this section we work out the effects of decoherence via
+collective-spin jump operators on ordered products of collective spin
+operators.  For shorthand, we define
+\begin{align}
+  \D\p{\Gamma;\O}
+  \equiv \D\p{\set{\Gamma};\O}
+  = \Gamma^\dag \O \Gamma - \f12\sp{\Gamma^\dag \Gamma, \O}_+,
+\end{align}
+where $\Gamma$ is a collective spin operator.
+
+\subsection{Decay-type decoherence and dephasing}
+\label{sec:decay_dephasing_collective}
+
+Making use of the results in Appendix \ref{sec:sandwich_collective},
+we find that the effects of collective decay-type decoherence on
+ordered products of collective spin operators are given by
+\begin{align}
+  \D\p{S_\bmu; S_\mu^\ell S_\z^m S_\bmu^n}
+  &= -S_\mu^{\ell+1} \p{\sp{\mu+S_\z}^m - S_\z^m} S_\bmu^{n+1} \notag \\
+  &\qquad + \f12 S_\mu^\ell \p{\ell\sp{\ell-1} + n\sp{n-1}
+    + 2\mu\sp{\ell+n} S_\z} S_\z^m S_\bmu^n,
+\end{align}
+and
+\begin{align}
+  \D\p{S_\mu; S_\mu^\ell S_\z^m S_\bmu^n}
+  &= S_\mu^{\ell+1} \p{\sp{2\mu + S_\z}^m-\sp{\mu + S_\z}^m}
+  S_\bmu^{n+1} \notag \\
+  &\qquad - S_\mu^\ell
+  \p{\ell\sp{\ell+1} + n\sp{n+1}+2\mu\sp{\ell+n+1} S_\z}
+  \p{\mu+S_\z}^m S_\bmu^n \notag \\
+  &\qquad + \f12 S_\mu^\ell
+  \p{\ell\sp{\ell+1} + n\sp{n+1}+2\mu\sp{\ell+n+2} S_\z}
+  S_\z^m S_\bmu^n \notag \\
+  &\qquad + \ell n S_\mu^{\ell-1}
+  \p{\sp{\ell-1}\sp{n-1}+2\mu\sp{\ell+n-2} S_\z + 4 S_\z^2}
+  S_\z^m S_\bmu^{n-1}.
+\end{align}
+Similarly, we can work out that the effect of collective spin
+dephasing is given by
+\begin{align}
+  \D\p{S_\z; S_\mu^\ell S_\z^m S_\bmu^n}
+  = -\f12 \p{\ell-n}^2 S_\mu^\ell S_\z^m S_\bmu^n.
+\end{align}
+
+\subsection{The general case}
+\label{sec:general_collective}
+
+More generally, we may need to consider collective-spin jump operators
+of the form
+\begin{align}
+  \Gamma \equiv \Gamma_\z S_\z + \Gamma_+ S_+ + \Gamma_- S_-,
+\end{align}
+whose decoherence effects are determined by
+\begin{align}
+  \Gamma^\dag \O \Gamma
+  = \abs{\Gamma_\z}^2 S_\z \O S_\z
+  + \sum_\mu \p{\abs{\Gamma_\mu}^2 S_\bmu \O S_\mu
+    + \Gamma_\bmu^* \Gamma_\mu S_\mu \O S_\mu
+    + \Gamma_\z^* \Gamma_\mu S_\z \O S_\mu
+    + \Gamma_\bmu^* \Gamma_\z S_\mu \O S_\z},
+\end{align}
+and
+\begin{align}
+  \Gamma^\dag \Gamma
+  = \abs{\Gamma_\z}^2 S_\z^2
+  + \sum_\mu \p{\abs{\Gamma_\mu}^2 S_\bmu S_\mu
+    + \Gamma_z^*\Gamma_\mu S_\z S_\mu
+    + \Gamma_\bmu^* \Gamma_z S_\mu S_\z
+    + \Gamma_\bmu^* \Gamma_\mu S_\mu^2},
 \end{align}
@@
 \begin{align}
-  \f{\mu}{2} \sp{S_z, S_\bmu^\ell S_z^m S_\mu^n}_+
-  = S_\bmu^\ell\p{-\f12\sp{\ell+n} + \mu S_z} S_z^m S_\mu^n,
-\end{align}
-and in turn
-\begin{align}
-  \D_\mu\p{S_\bmu^\ell S_z^m S_\mu^n}
-  &= S_\bmu^\ell \sp{\p{S+S_z} \p{\mu+S_z}^m
-    - \p{S + \f12\sp{\ell+n} + S_z} S_z^m} S_\mu^n \\
-  &= S_\bmu^\ell
-  \sp{\p{S+S_z} \sum_{k=0}^{m-1} \mu^{m-k} { m \choose k} S_z^k
-  - \f12\p{\ell+n} S_z^m} S_\mu^n.
-\end{align}
-Decoherence through $\D_\mu$ only couples operators of the form
-$S_\bmu^\ell S_z^m S_\mu^n$ to operators of the (almost identical)
-form $S_\bmu^\ell S_z^{m'} S_\mu^n$ with $m'\le m$.
-
-\subsection{The harder case}
-
-The second case to consider for the decoherence of general ordered
-collective spin operators is
-\begin{align}
-  \D_\mu\p{S_\mu^\ell S_z^m S_\bmu^n}
-  = \sum_j\sigma_j^\bmu S_\mu^\ell S_z^m S_\bmu^n \sigma_j^\mu
-  - S S_\mu^\ell S_z^m S_\bmu^n
-  + \f{\mu}{2} \sp{S_z, S_\mu^\ell S_z^m S_\bmu^n}_+.
-  \label{eq:D_mu_mu}
-\end{align}
-The difficulty now is with the first term in \eqref{eq:D_mu_mu}, as
-$\sigma_j^\bmu$ and $S_\mu$ do not commute.
+  \D\p{\Gamma;\O}
+  &= \sum_{X\in\set{\z,+,-}} \abs{\Gamma_X}^2 \D\p{S_X;\O}
+  + \sum_\mu \p{\Gamma_\bmu^* \Gamma_\mu S_\mu \O S_\mu
+    + \Gamma_\z^* \Gamma_\mu S_\z \O S_\mu
+    + \Gamma_\bmu^* \Gamma_\z S_\mu \O S_\z}
+  \notag \\
+  &\qquad -\f12 \sum_\mu\p{\Gamma_z^*\Gamma_\mu \sp{S_\z S_\mu, \O}_+
+    + \Gamma_\bmu^* \Gamma_z \sp{S_\mu S_\z, \O}_+
+    + \Gamma_\bmu^* \Gamma_\mu \sp{S_\mu^2, \O}_+}.
+  \label{eq:D_general_collective}
+\end{align}
+In order to compute the effect of this decoherence on ordered products
+of collective spin operators, we expand the anti-commutators
+\begin{align}
+  \sp{S_\z S_\mu, S_\mu^\ell S_\z^m S_\bmu^n}_+
+  &= S_\mu^{\ell+1} \sp{\p{\mu\sp{\ell+1} + S_\z} S_\z^m
+    + \p{\mu\sp{n+1} + S_\z} \p{\mu+S_\z}^m } S_\bmu^n \notag \\
+  &\qquad - n S_\mu^\ell \p{\mu n \sp{n-1}
+    + \sp{3n-1} S_\z + 2\mu S_\z^2} S_\z^m S_\bmu^{n-1},
+  \allowdisplaybreaks \\
+  \sp{S_\bmu S_\z, S_\mu^\ell S_\z^m S_\bmu^n}_+
+  &= S_\mu^\ell \sp{\p{\mu\sp{\ell+1} + S_\z} \p{\mu+S_\z}^m
+    + \p{\mu\sp{n+1} + S_\z} S_\z^m } S_\bmu^{n+1} \notag \\
+  &\qquad - \ell S_\mu^{\ell-1} \p{\mu \ell \sp{\ell-1}
+    + \sp{3\ell-1} S_\z + 2\mu S_\z^2} S_\z^m S_\bmu^n,
+  \allowdisplaybreaks \\
+  \sp{S_\mu S_\z, S_\mu^\ell S_\z^m S_\bmu^n}_+
+  &= S_\mu^{\ell+1} \sp{\p{\mu\ell+S_\z} S_\z^m
+    + \p{\mu n+S_\z} \p{\mu+S_\z}^m} S_\bmu^n \notag \\
+  &\qquad - n S_\mu^\ell \p{\mu\sp{n-1}^2
+    + 3\sp{n-1} S_\z + 2\mu S_\z^2} S_\z^m S_\bmu^{n-1},
+  \allowdisplaybreaks \\
+  \sp{S_\z S_\bmu, S_\mu^\ell S_\z^m S_\bmu^n}_+
+  &= S_\mu^\ell \sp{\p{\mu\ell+S_\z} \p{\mu+S_\z}^m
+    + \p{\mu n+S_\z} S_\z^m} S_\bmu^{n+1} \notag \\
+  &\qquad - \ell S_\mu^{\ell-1} \p{\mu\sp{\ell-1}^2
+    + 3\sp{\ell-1} S_\z + 2\mu S_\z^2} S_\z^m S_\bmu^n,
+  \allowdisplaybreaks \\
+  \sp{S_\mu^2, S_\mu^\ell S_\z^m S_\bmu^n}_+
+  &= S_\mu^{\ell+2} \p{S_\z^m+\sp{2\mu+S_\z}^m} S_\bmu^n
+  - 2n S_\mu^{\ell+1} \p{n+2\mu S_\z} \p{\mu+S_\z}^m S_\bmu^{n-1}
+  \notag \\
+  &\qquad + n\p{n-1} S_\mu^\ell \p{\sp{n-1}\sp{n-2}
+    + 2\mu\sp{2n-3} S_\z + 4 S_\z^2} S_\z^m S_\bmu^{n-2},
+  \allowdisplaybreaks \\
+  \sp{S_\bmu^2, S_\mu^\ell S_\z^m S_\bmu^n}_+
+  &= S_\mu^\ell \p{S_\z^m+\sp{2\mu+S_\z}^m} S_\bmu^{n+2}
+  - 2\ell S_\mu^{\ell-1} \p{\ell+2\mu S_\z} \p{\mu+S_\z}^m S_\bmu^{n+1}
+  \notag \\
+  &\qquad + \ell\p{\ell-1} S_\mu^{\ell-2} \p{\sp{\ell-1}\sp{\ell-2}
+    + 2\mu\sp{2\ell-3} S_\z + 4 S_\z^2} S_\z^m S_\bmu^n.
+\end{align}
+As with the case of general single-spin decoherence, there is not much
+to gain from substituting $\O = S_\mu^\ell S_\z^m S_\bmu^n$ into
+\eqref{eq:D_general_collective} and further simplifying the resulting
+expression; the anti-commutators provided in above together with the
+results in in Sections \ref{sec:sandwich_collective} and
+\ref{sec:decay_dephasing_collective} are sufficient to compute the
+effects of collective-spin decoherence when necessary.
 
@@
 \begin{align}
-  S_z^n = \sum_{k=-S}^S k^n \op{k},
-  &&
-  S_\mu
-  = \sum_{k=-S+\delta_{\mu,-1}}^{S-\delta_{k,1}}
-  g_\mu\p{k} \op{k+\mu}{k},
+  S_\z = \sum_{k=-S}^S k \op{k},
+  &&
+  S_\mu = \sum_{k=-S+\delta_{\mu,-1}}^{S-\delta_{\mu,1}}
+  g_\mu\p{S,k} \op{k+\mu}{k},
 \end{align}
@@
 \begin{align}
-  \ket{\pm Z} \equiv \ket{\pm S},
-  &&
-  \ket{X} \equiv 2^{-N/2} \sum_{k=-S}^S { N \choose S+k }^{1/2} \ket{k},
-\end{align}
-using which we can compute
-\begin{align}
-  \bk{\pm Z|S_+^\ell S_z^m S_-^n|\pm Z}
+  \ket{\pm\Z} \equiv \ket{\pm S},
+  &&
+  \ket{\X} \equiv 2^{-N/2} \sum_{k=-S}^S
+  { N \choose S+k }^{1/2} \ket{k},
+\end{align}
+using which we can compute, for $\bmu\equiv-\mu\in\set{+1,-1}$,
+\begin{align}
+  \bk{\pm\Z|S_\mu^\ell S_\z^m S_\bmu^n|\pm\Z}
   = \delta_{\ell,0} \delta_{m,0} \p{\pm S}^m,
@@
 \begin{align}
-  \bk{X|S_+^\ell S_z^m S_-^n|X}
+  \bk{\X|S_+^\ell S_\z^m S_-^n|\X}
   &=  2^{-N} \sum_{k=-S}^{S-\max\set{\ell,n}} k^m
-  \sp{{ N \choose S+k+\ell } { N \choose S+k+n }}^{1/2} \tag*{} \\
+  \sp{{ N \choose S+k+\ell } { N \choose S+k+n }}^{1/2} \notag \\
   &\qquad\qquad\qquad\qquad
@@
 \end{align}
-With the help of Mathematica, we can write the sum in the alternate
-form
-\begin{align}
-  \bk{X|S_+^\ell S_z^m S_-^n|X}
+With the help of Mathematica, we can write this result in the
+alternate form
+\begin{align}
+  \bk{\X|S_+^\ell S_\z^m S_-^n|\X}
   = \f{\p{N!}^2}{2^N\p{N-\ell}!\p{N-n}!}
@@
   \equiv \F\p{\let\scriptstyle\textstyle
-    \substack{\ell-N,~n-N\\-N}~; -1},
+    \substack{\ell-N,~n-N \\ -N}~; -1},
 \end{align}
@@
   \tilde\F_{m-p-1}\p{\let\scriptstyle\textstyle
-    \substack{1+\ell-N,~1+n-N\\1-N}~; -1},
+    \substack{1+\ell-N,~1+n-N \\ 1-N}~; -1},
 \end{align}
@@
   \equiv \F\p{\let\scriptstyle\textstyle
-    \substack{2,2,\cdots,2,a,b\\1,1,\cdots,1,c}~; z}
+    \substack{2,2,\cdots,2,a,b \\ 1,1,\cdots,1,c}~; z}
 \end{align}

### 2018-09-06 soc_squeezing/cumulant_expansions.tex
added initial conditions for (mu,z,bmu), rather than just (+,z,-) removed redundand sandwich identities changed OAT correlators to incorporate arbitrary single-spin decoherence

@@
 expectation values with respect to the initial state $\ket{\X}$
-satisfying $S_\x\ket{\X}=S\ket{\X}$, in terms of
+satisfying $S_\x\ket{\X}=S\ket{\X}$, in terms of $S\equiv N/2$ and
 $\mu,\nu\in\set{+1,-1}$ we find that
@@
   &= \sum_j\bk{\sigma_+^{(j)}}
-  = \f{N}{2} e^{-\gamma t/2} \Phi\p{\chi,t}^{N-1}, \label{eq:S+_OAT} \\
+  = S e^{-\Gamma t} \Phi\p{\chi,t}^{N-1}, \label{eq:S+_OAT} \\
   \bk{S_\mu S_\z}
@@
   + \f12\sum_{j\ne k} \bk{\sigma_\mu^{(j)} \sigma_\z^{(k)}} \\
-  &= -\f{\mu}{2}\bk{S_\mu} + \f14 N \p{N-1} e^{-\gamma t/2}
+  &= -\f{\mu}{2}\bk{S_\mu} + S \p{S-\f12} e^{-\Gamma t}
   \Psi\p{\mu\chi,t} \Phi\p{\chi,t}^{N-2}, \\
@@
   + \sum_{j\ne k} \bk{\sigma_\mu^{(j)} \sigma_\nu^{(k)}} \\
-  &= \delta_{\mu,-\nu} \p{\f{N}{2} + \mu \bk{S_\z}}
-  + \f14 N \p{N-1} e^{-\gamma t}
+  &= \delta_{\mu,-\nu} \p{S + \mu\bk{S_\z}}
+  + S \p{S-\f12} e^{-2\Gamma t}
   \Phi\p{\sp{\mu+\nu}\chi,t}^{N-2}, \label{eq:SS+-_OAT}
@@
 \begin{align}
-  \Phi\p{\zeta,t} \equiv e^{-\gamma t/2}
-  \sp{\cos\p{s_\zeta t}+\f12\gamma t~\t{sinc}\p{s_\zeta t}},
-  &&
-  \Psi\p{\zeta,t} \equiv e^{-\gamma t/2}
-  \p{is_\zeta-\gamma/2}t~ \t{sinc}\p{s_\zeta t},
-\end{align}
-and $s_\zeta \equiv \zeta + i2\gamma$.  The last ingredients we need
-to compute spin squeezing at any time are $\bk{S_\z}$ and
-$\bk{S_\z^2}$.  As these operators commute with the OAT Hamiltonian,
-their evolution is governed entirely by decoherence through
-uncorrelated decay of individual spins, which means (see Appendix
-\ref{sec:decay_single})
-\begin{align}
-  \f{d}{dt} S_\z = -\gamma \p{S + S_\z},
-  &&
-  \f{d}{dt}\p{S_\z^2} = \gamma \p{S + S_\z} \p{1 - 2S_\z},
-\end{align}
-where $S\equiv N/2$.  The initial conditions $\bk{S_\z}_{t=0}=0$ and
-$\bk{S_\z^2}_{t=0}=S/2$ then imply
-\begin{align}
-  \bk{S_\z} = S\p{e^{-\gamma t}-1},
-  &&
-  \bk{S_\z^2} = S \p{1 - \f12 e^{-\gamma t}} e^{-\gamma t} + \bk{S_\z}^2.
+  \Phi\p{\zeta,t}
+  &\equiv e^{-\lambda t} \sp{\cos\p{t\sqrt{s_\zeta^2-r}}
+    + \lambda t~\t{sinc}\p{t\sqrt{s_\zeta^2-r}}},
+  \\
+  \Psi\p{\zeta,t}
+  &\equiv e^{-\lambda t} \p{is_\zeta-\gamma}t~
+  \t{sinc}\p{t\sqrt{s_\zeta^2-r}},
+\end{align}
+for
+\begin{align}
+  \gamma \equiv \f12 \p{\gamma_+ - \gamma_-},
+  &&
+  \lambda \equiv \f12 \p{\gamma_+ + \gamma_-},
+  &&
+  r \equiv \gamma_+ \gamma_-,
+  &
+  \Gamma \equiv \gamma_z + \lambda,
+  &&
+  s_\zeta \equiv \zeta + i\gamma,
+\end{align}
+defined respectively in terms of spontaneous uncorrelated spin
+dephasing, excitation, and decay rates $\gamma_z$, $\gamma_+$,
+$\gamma_-$.  The last ingredients we need to compute spin squeezing at
+any time are $\bk{S_\z}$ and $\bk{S_\z^2}$.  As these operators
+commute with the OAT Hamiltonian, their evolution is governed entirely
+by decoherence (see Appendix \ref{sec:decay_single}), which means
+\begin{align}
+  \f{d}{dt} S_\z
+  &= S\p{\gamma_+-\gamma_-} - \p{\gamma_++\gamma_-} S_\z,
+  \\
+  \f{d}{dt}\p{S_\z^2}
+  &= S\p{\gamma_++\gamma_-} + 2\p{\gamma_+-\gamma_-}\p{S-\f12} S_\z
+  - 2 \p{\gamma_++\gamma_-} S_\z^2.
+\end{align}
+The initial conditions $\bk{S_\z}_{t=0}=0$ and $\bk{S_\z^2}_{t=0}=S/2$
+then imply
+\begin{align}
+  \bk{S_\z}
+  = S\p{\f{\gamma_+-\gamma_-}{\gamma_++\gamma_-}}
+  \p{1-e^{-\p{\gamma_+ + \gamma_-} t}},
+  &&
+  \bk{S_\z^2} = S \sp{\f12 + \p{S-\f12} \f{\bk{S_z}^2}{S^2}}.
   \label{eq:Sz_OAT}
@@
 
-\section{Truncated Heisenberg operator dynamics}
+\section{Truncated operator dynamics}
 
@@
 
+\section{One-axis twisting revisited}
+
 \bibliography{\jobname}
@@
 \end{align}
-Working through the combinations, we then compute
-\begin{align}
-  \sum_j \sigma_\mu^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\bmu^{(j)}
-  &= S_\mu^\ell \p{S+\mu S_\z}\p{-\mu+S_\z}^m S_\bmu^n,
-  \label{eq:sand_mu_bmu} \allowdisplaybreaks \\
-  \sum_j \sigma_\mu^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\z^{(j)}
-  &= -\mu S_\mu^{\ell+1} S_\z^m S_\bmu^n
-  + 2\mu n S_\mu^\ell \p{S+\mu S_\z} \p{-\mu+S_\z}^m S_\bmu^{n-1},
-  \allowdisplaybreaks \\
-  \sum_j \sigma_\z^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\bmu^{(j)}
-  &= -\mu S_\mu^\ell S_\z^m S_\bmu^{n+1}
-  + 2\mu \ell S_\mu^{\ell-1} \p{S+\mu S_\z}\p{-\mu+S_\z}^m S_\bmu^n,
-  \allowdisplaybreaks \\
+Working through all combinations that are unique up to conjugation, we
+then compute
+\begin{align}
   \sum_j \sigma_\z^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\z^{(j)}
@@
   \label{eq:sand_z_z} \allowdisplaybreaks \\
-  \sum_j \sigma_\mu^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\mu^{(j)}
-  &= -n\p{n-1} S_\mu^\ell \p{S+\mu S_\z} \p{-\mu+S_\z}^m S_\bmu^{n-2}
-  + n S_\mu^{\ell+1} S_\z^m S_\bmu^{n-1},
-  \allowdisplaybreaks \\
-  \sum_j \sigma_\bmu^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\bmu^{(j)}
-  &= -\ell \p{\ell-1} S_\mu^{\ell-2}
-  \p{S+\mu S_\z} \p{-\mu+S_\z}^m S_\bmu^n
-  + \ell S_\mu^{\ell-1} S_\z^m S_\bmu^{n+1},
-  \allowdisplaybreaks \\
   \sum_j \sigma_\z^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\mu^{(j)}
@@
   \p{S+\mu S_\z} \p{-\mu+S_\z}^m S_\bmu^{n-2},
-  \allowdisplaybreaks \\
-  \sum_j \sigma_\bmu^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\z^{(j)}
-  &= \mu S_\mu^\ell \p{\mu+S_\z}^m S_\bmu^{n+1}
-  + \mu \ell \p{-2S+2n+\ell-1} S_\mu^{\ell-1} S_\z^m S_\bmu^n \notag \\
-  &\qquad - 2\mu\ell n\p{\ell-1} S_\mu^{\ell-2}
-  \p{S+\mu S_\z} \p{-\mu+S_\z}^m S_\bmu^{n-1},
-  \allowdisplaybreaks \\
+  \label{eq:sand_z_mu} \allowdisplaybreaks \\
+  \sum_j \sigma_\z^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\bmu^{(j)}
+  &= -\mu S_\mu^\ell S_\z^m S_\bmu^{n+1}
+  + 2\mu \ell S_\mu^{\ell-1} \p{S+\mu S_\z}\p{-\mu+S_\z}^m S_\bmu^n,
+  \label{eq:sand_z_bmu} \allowdisplaybreaks \\
+  \sum_j \sigma_\mu^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\mu^{(j)}
+  &= -n\p{n-1} S_\mu^\ell \p{S+\mu S_\z} \p{-\mu+S_\z}^m S_\bmu^{n-2}
+  + n S_\mu^{\ell+1} S_\z^m S_\bmu^{n-1},
+  \label{eq:sand_mu_mu} \allowdisplaybreaks \\
+  \sum_j \sigma_\mu^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\bmu^{(j)}
+  &= S_\mu^\ell \p{S+\mu S_\z}\p{-\mu+S_\z}^m S_\bmu^n,
+  \label{eq:sand_mu_bmu} \allowdisplaybreaks \\
   \sum_j \sigma_\bmu^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\mu^{(j)}
@@
 
+\subsection{Dephasing}
+\label{sec:dephasing_single}
+
+The effect of decoherence via single-spin dephasing is described by
+\begin{align}
+  \D\p{\sigma_\z/\sqrt2; \O}
+  = \sum_j\p{\sigma_\z^{(j)} \O \sigma_\z^{(j)}
+    - \f12\sp{\sigma_\z^{(j)} \sigma_\z^{(j)},\O}_+}
+  = \f12 \sum_j \sigma_\z^{(j)} \O \sigma_\z^{(j)} - S \O,
+\end{align}
+where the factor $\sqrt2$ is inserted as
+e.g.~$\D\p{\sigma_\z/\sqrt2;S_\x}=-S_\x$.  From \eqref{eq:sand_z_z},
+we then have
+\begin{align}
+  \D\p{\sigma_\z/\sqrt2; S_\mu^\ell S_\z^m S_\bmu^n}
+  = -\p{\ell+n} S_\mu^\ell S_\z^m S_\bmu^n
+  + 2 \ell n S_\mu^{\ell-1} \p{S + \mu S_\z}
+  \p{-\mu + S_\z}^m S_\bmu^{n-1}.
+\end{align}
+Decoherence via single-spin dephasing makes operators
+$S_\mu^\ell S_\z^m S_\bmu^n$ ``grow'' in $S_\z$, although the sum
+$\ell+m+n$ does not grow.
+
 \subsection{Decay-type decoherence}
@@
   \D\p{\sigma_\bmu; S_\mu^\ell S_\z^m S_\bmu^n}
-  &= S_\mu^\ell \p{S+\mu S_\z}\p{-\mu+S_\z}^m S_\bmu^n
-  - S_\mu^\ell\p{S + \f12\sp{\ell+n} + \mu S_\z} S_\z^m S_\bmu^n \\
-  &= S_\mu^\ell \sp{\p{S+\mu S_\z}
-    \sum_{k=0}^{m-1} \p{-\mu}^{m-k} { m \choose k } S_\z^k
-    - \f12 \p{\ell+n} S_\z^m} S_\bmu^n,
+  = S_\mu^\ell \p{S+\mu S_\z}\p{-\mu+S_\z}^m S_\bmu^n
+  - S_\mu^\ell\p{S + \f12\sp{\ell+n} + \mu S_\z} S_\z^m S_\bmu^n,
   \label{eq:decay_diff}
@@
   &= S_\mu^\ell \p{S - \ell - n - \mu S_\z} \p{\mu + S_\z}^m S_\bmu^n
-  + \ell n \p{2S - \ell - n + 2}
-  S_\mu^{\ell-1} S_\z^m S_\bmu^{n-1} \notag \\
-  &\qquad + \ell n \p{\ell-1} \p{n-1} S_\mu^{\ell-2} \p{S + \mu S_\z}
-  \p{-\mu + S_\z}^m S_\bmu^{n-2} \notag \\
-  &\qquad - S_\mu^\ell\p{S - \f12\sp{\ell+n} - \mu S_\z}
-  S_\z^m S_\bmu^n \allowdisplaybreaks \\
-  &= S_\mu^\ell \sp{\p{S-\ell-n-\mu S_\z}
-    \sum_{k=0}^{m-1} \mu^{m-k} { m \choose k } S_\z^k
-    - \f12 \p{\ell+n} S_\z^m} S_\bmu^n \notag \\
+  - S_\mu^\ell\p{S - \f12\sp{\ell+n} - \mu S_\z}
+  S_\z^m S_\bmu^n \notag \\
   &\qquad + \ell n \p{2S - \ell - n + 2}
@@
   &\qquad + \ell n \p{\ell-1} \p{n-1} S_\mu^{\ell-2} \p{S + \mu S_\z}
-  \sum_{k=0}^m \p{-\mu}^{m-k} { m \choose k } S_\z^k S_\bmu^{n-2}.
+  \p{-\mu + S_\z}^m S_\bmu^{n-2}.
   \label{eq:decay_same}
@@
 not grow.
-
-\subsection{Dephasing}
-\label{sec:dephasing_single}
-
-The effect of decoherence via single-spin dephasing is described by
-\begin{align}
-  \D\p{\sigma_\z/\sqrt2; \O}
-  = \sum_j\p{\sigma_\z^{(j)} \O \sigma_\z^{(j)}
-    - \f12\sp{\sigma_\z^{(j)} \sigma_\z^{(j)},\O}_+}
-  = \f12 \sum_j \sigma_\z^{(j)} \O \sigma_\z^{(j)} - S \O,
-\end{align}
-where the factor $\sqrt2$ is inserted as
-e.g.~$\D\p{\sigma_\z/\sqrt2;S_\x}=-S_\x$.  From \eqref{eq:sand_z_z},
-we then have
-\begin{align}
-  \D\p{\sigma_\z/\sqrt2; S_\mu^\ell S_\z^m S_\bmu^n}
-  &= -\p{\ell+n} S_\mu^\ell S_\z^m S_\bmu^n
-  + 2 \ell n S_\mu^{\ell-1} \p{S + \mu S_\z}
-  \p{-\mu + S_\z}^m S_\bmu^{n-1} \\
-  &= -\p{\ell+n} S_\mu^\ell S_\z^m S_\bmu^n
-  + 2 \ell n S_\mu^{\ell-1} \p{S + \mu S_\z}
-  \sum_{k=0}^m \p{-\mu}^{m-k} { m \choose k } S_\z^k S_\bmu^{n-1}.
-\end{align}
-Decoherence via single-spin dephasing also makes operators
-$S_\mu^\ell S_\z^m S_\bmu^n$ ``grow'' in $S_\z$, although the sum
-$\ell+m+n$ still does not grow.
 
@@
 In order to compute the effect of this decoherence on ordered products
-of collective spin operators, we expand the anti-commutators
+of collective spin operators, we expand the anti-commutator
 \begin{align}
@@
   &= S_\mu^{\ell+1} \sp{S_\z^m+\p{\mu+S_\z}^m} S_\bmu^n
-  - n S_\mu^\ell \p{n-1+2\mu S_\z} S_\z^m S_\bmu^{n-1}, \\
-  \sp{S_\bmu, S_\mu^\ell S_\z^m S_\bmu^n}_+
-  &= S_\mu^\ell \sp{S_\z^m+\p{\mu+S_\z}^m} S_\bmu^{n+1}
-  - \ell S_\mu^{\ell-1} \p{\ell-1+2\mu S_\z} S_\z^m S_\bmu^n.
+  - n S_\mu^\ell \p{n-1+2\mu S_\z} S_\z^m S_\bmu^{n-1}.
   \label{eq:S_mu_acomm}
 \end{align}
-These anti-commutators together with the results in Sections
+This anti-commutator together with the results in Sections
 \ref{sec:sandwich_single}, \ref{sec:decay_single}, and
@@
   &= S_\mu^{\ell+1} \p{\mu n+S_\z} S_\z^m S_\bmu^n,
-  \allowdisplaybreaks \\
-  S_\z S_\mu^\ell S_\z^m S_\bmu^n S_\bmu
-  &= S_\mu^\ell \p{\mu\ell+S_\z} S_\z^m S_\bmu^{n+1},
   \allowdisplaybreaks \\
@@
   \allowdisplaybreaks \\
-  S_\bmu S_\mu^\ell S_\z^m S_\bmu^n S_\bmu
-  &= S_\mu^\ell \p{\mu+S_\z}^m S_\bmu^{n+2}
-  - \ell S_\mu^{\ell-1} \p{\ell-1+2\mu S_\z} S_\z^m S_\bmu^{n+1},
-  \allowdisplaybreaks \\
   S_\z S_\mu^\ell S_\z^m S_\bmu^n S_\mu
@@
   \p{\mu\ell\sp{n-1} + \sp{2\ell+n-1}S_\z + 2\mu S_\z^2}
-  S_\z^m S_\bmu^{n-1},
-  \allowdisplaybreaks \\
-  S_\bmu S_\mu^\ell S_\z^m S_\bmu^n S_\z
-  &= S_\mu^\ell \p{\mu\sp{n+1}+S_\z} \p{\mu+S_\z}^m S_\bmu^{n+1}
-  \notag \\
-  &\qquad - \ell S_\mu^{\ell-1}
-  \p{\mu n\sp{\ell-1} + \sp{2n+\ell-1}S_\z + 2\mu S_\z^2}
-  S_\z^m S_\bmu^n.
+  S_\z^m S_\bmu^{n-1}.
 \end{align}
@@
 \begin{align}
-  \bk{\X|S_+^\ell S_\z^m S_-^n|\X}
-  &=  2^{-N} \sum_{k=-S}^{S-\max\set{\ell,n}} k^m
-  \sp{{ N \choose S+k+\ell } { N \choose S+k+n }}^{1/2} \notag \\
+  \bk{\X|S_\mu^\ell S_\z^m S_\bmu^n|\X}
+  &=  2^{-N} \sum_{k=-S+\delta_{\mu,-1}\max\set{\ell,n}}
+  ^{S-\delta_{\mu,1}\max\set{\ell,n}} k^m
+  \sp{{ N \choose S+k+\mu\ell } { N \choose S+k+\mu n }}^{1/2}
+  \notag \\
   &\qquad\qquad\qquad\qquad
-  \times \prod_{p=0}^{\ell-1} g_+\p{S,k+p}
-  \prod_{q=0}^{n-1} g_-\p{S,k+n-q} \\
-  &= \f{N!}{2^N} \sum_{k=-S}^{S-\max\set{\ell,n}} k^m
-  \f{\p{S-k}!}{\p{S+k}!\p{S-k-\ell}!\p{S-k-n}!}.
-\end{align}
-With the help of Mathematica, we can write this result in the
-alternate form
-\begin{align}
-  \bk{\X|S_+^\ell S_\z^m S_-^n|\X}
-  = \f{\p{N!}^2}{2^N\p{N-\ell}!\p{N-n}!}
-  \sp{\p{-S}^m \alpha + \f{\p{N-\ell}\p{N-n}}{N} \beta},
-\end{align}
-where $\alpha$ and $\beta$ are defined in terms of the generalized
-hypergeometric function $\F$ by
-\begin{align}
-  \alpha
-  \equiv \F\p{\let\scriptstyle\textstyle
-    \substack{\ell-N,~n-N \\ -N}~; -1},
-\end{align}
-and
-\begin{align}
-  \beta
-  \equiv \sum_{p=0}^{m-1} \p{-S}^p { m \choose p }
-  \tilde\F_{m-p-1}\p{\let\scriptstyle\textstyle
-    \substack{1+\ell-N,~1+n-N \\ 1-N}~; -1},
-\end{align}
-for
-\begin{align}
-  \tilde\F_q\p{\let\scriptstyle\textstyle
-    \substack{a,~b\\c}~; z}
-  \equiv \F\p{\let\scriptstyle\textstyle
-    \substack{2,2,\cdots,2,a,b \\ 1,1,\cdots,1,c}~; z}
-\end{align}
-with $q$ repetitions of 2 and 1.
+  \times \prod_{p=0}^{\ell-1} g_\mu\p{S,k+\mu p}
+  \prod_{q=0}^{n-1} g_\bmu\p{S,k+\mu\sp{n-q}} \\
+  &= \f{N!}{2^N} \sum_{k=-S+\delta_{\mu,-1}\max\set{\ell,n}}
+  ^{S-\delta_{\mu,1}\max\set{\ell,n}} k^m
+  \f{\p{S-\mu k}!}{\p{S+\mu k}!\p{S-\mu k-\ell}!\p{S-\mu k-n}!}.
+\end{align}
+% With the help of Mathematica, we can write this result in the
+% alternate form
+% \begin{align}
+%   \bk{\X|S_+^\ell S_\z^m S_-^n|\X}
+%   = \f{\p{N!}^2}{2^N\p{N-\ell}!\p{N-n}!}
+%   \sp{\p{-S}^m \alpha + \f{\p{N-\ell}\p{N-n}}{N} \beta},
+% \end{align}
+% where $\alpha$ and $\beta$ are defined in terms of the generalized
+% hypergeometric function $\F$ by
+% \begin{align}
+%   \alpha
+%   \equiv \F\p{\let\scriptstyle\textstyle
+%     \substack{\ell-N,~n-N \\ -N}~; -1},
+% \end{align}
+% and
+% \begin{align}
+%   \beta
+%   \equiv \sum_{p=0}^{m-1} \p{-S}^p { m \choose p }
+%   \tilde\F_{m-p-1}\p{\let\scriptstyle\textstyle
+%     \substack{1+\ell-N,~1+n-N \\ 1-N}~; -1},
+% \end{align}
+% for
+% \begin{align}
+%   \tilde\F_q\p{\let\scriptstyle\textstyle
+%     \substack{a,~b\\c}~; z}
+%   \equiv \F\p{\let\scriptstyle\textstyle
+%     \substack{2,2,\cdots,2,a,b \\ 1,1,\cdots,1,c}~; z}
+% \end{align}
+% with $q$ repetitions of 2 and 1.
 

### 2018-09-07 soc_squeezing/cumulant_expansions.tex
drastically simplified general decoherence results

@@
   &&
-  \bk{S_\z^2} = S \sp{\f12 + \p{S-\f12} \f{\bk{S_z}^2}{S^2}}.
+  \bk{S_\z^2} = S \sp{\f12 + \p{S-\f12} \f{\bk{S_\z}^2}{S^2}}.
   \label{eq:Sz_OAT}
@@
   &= \mu \sum_{k=0}^{mq-1} S_\mu^k
-  \p{2\mu\sp{m-k-1} \sigma_\mu^{(j)} S_\mu^{m-k-2}
+  \sp{2\mu\p{m-k-1} \sigma_\mu^{(j)} S_\mu^{m-k-2}
     + S_\mu^{m-k-1} \sigma_\z^{(j)}} \\
@@
 We are now equipped to derive similar identities for general ordered
-products of collective spin operators.  For quick reference, we
-provide here identity \eqref{eq:push_z_mu_sS} and its conjugate:
-\begin{align}
-  \sigma_\z^{(j)} S_\mu^\ell
-  = S_\mu^\ell \sigma_\z^{(j)} + 2\mu \ell S_\mu^{\ell-1}\sigma_\mu^{(j)},
-  &&
-  S_\bmu^n \sigma_\z^{(j)}
-  = \sigma_\z^{(j)} S_\bmu^n + 2\mu n \sigma_\bmu^{(j)} S_\bmu^{n-1},
-\end{align}
-and likewise with identity \eqref{eq:push_mu_Ss}:
-\begin{align}
-  \sigma_\bmu^{(j)} S_\mu^\ell
-  &= S_\mu^\ell \sigma_\bmu^{(j)}
-  - \ell \p{\ell-1} S_\mu^{\ell-2} \sigma_\mu^{(j)}
-  - \mu \ell S_\mu^{\ell-1} \sigma_\z^{(j)}, \\
-  S_\bmu^n \sigma_\mu^{(j)}
-  &= \sigma_\mu^{(j)} S_\bmu^n
-  - n \p{n-1} \sigma_\bmu^{(j)} S_\bmu^{n-2}
-  - \mu n \sigma_\z^{(j)} S_\bmu^{n-1}.
-\end{align}
-Working through all combinations that are unique up to conjugation, we
-then compute
+products of collective spin operators.  Making heavy use of identities
+\eqref{eq:push_z_mu_sS} and \eqref{eq:push_mu_Ss} to push single-spin
+operators through transverse collective-spin operators, we work
+through all combinations that are unique up to Hermitian conjugation,
+finding
 \begin{align}
@@
   \p{-\mu+S_\z}^m S_\bmu^{n-1},
-  \label{eq:sand_z_z} \allowdisplaybreaks \\
+  \label{eq:san_z_z} \allowdisplaybreaks \\
   \sum_j \sigma_\z^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\mu^{(j)}
@@
   \p{S+\mu S_\z} \p{-\mu+S_\z}^m S_\bmu^{n-2},
-  \label{eq:sand_z_mu} \allowdisplaybreaks \\
+  \label{eq:san_z_mu} \allowdisplaybreaks \\
   \sum_j \sigma_\z^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\bmu^{(j)}
@@
   + 2\mu \ell S_\mu^{\ell-1} \p{S+\mu S_\z}\p{-\mu+S_\z}^m S_\bmu^n,
-  \label{eq:sand_z_bmu} \allowdisplaybreaks \\
+  \label{eq:san_z_bmu} \allowdisplaybreaks \\
   \sum_j \sigma_\mu^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\mu^{(j)}
-  &= -n\p{n-1} S_\mu^\ell \p{S+\mu S_\z} \p{-\mu+S_\z}^m S_\bmu^{n-2}
-  + n S_\mu^{\ell+1} S_\z^m S_\bmu^{n-1},
-  \label{eq:sand_mu_mu} \allowdisplaybreaks \\
+  &= n S_\mu^{\ell+1} S_\z^m S_\bmu^{n-1}
+  - n\p{n-1} S_\mu^\ell \p{S+\mu S_\z} \p{-\mu+S_\z}^m S_\bmu^{n-2},
+  \label{eq:san_mu_mu} \allowdisplaybreaks \\
   \sum_j \sigma_\mu^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\bmu^{(j)}
   &= S_\mu^\ell \p{S+\mu S_\z}\p{-\mu+S_\z}^m S_\bmu^n,
-  \label{eq:sand_mu_bmu} \allowdisplaybreaks \\
+  \label{eq:san_mu_bmu} \allowdisplaybreaks \\
   \sum_j \sigma_\bmu^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\mu^{(j)}
@@
   \p{-\mu + S_\z}^m S_\bmu^{n-2}.
-  \label{eq:sand_bmu_mu}
+  \label{eq:san_bmu_mu}
 \end{align}
@@
 the identity on all other spins.
-
-\subsection{Dephasing}
-\label{sec:dephasing_single}
-
-The effect of decoherence via single-spin dephasing is described by
-\begin{align}
-  \D\p{\sigma_\z/\sqrt2; \O}
-  = \sum_j\p{\sigma_\z^{(j)} \O \sigma_\z^{(j)}
-    - \f12\sp{\sigma_\z^{(j)} \sigma_\z^{(j)},\O}_+}
-  = \f12 \sum_j \sigma_\z^{(j)} \O \sigma_\z^{(j)} - S \O,
-\end{align}
-where the factor $\sqrt2$ is inserted as
-e.g.~$\D\p{\sigma_\z/\sqrt2;S_\x}=-S_\x$.  From \eqref{eq:sand_z_z},
-we then have
-\begin{align}
-  \D\p{\sigma_\z/\sqrt2; S_\mu^\ell S_\z^m S_\bmu^n}
-  = -\p{\ell+n} S_\mu^\ell S_\z^m S_\bmu^n
-  + 2 \ell n S_\mu^{\ell-1} \p{S + \mu S_\z}
-  \p{-\mu + S_\z}^m S_\bmu^{n-1}.
-\end{align}
-Decoherence via single-spin dephasing makes operators
-$S_\mu^\ell S_\z^m S_\bmu^n$ ``grow'' in $S_\z$, although the sum
-$\ell+m+n$ does not grow.
 
@@
 \end{align}
-which implies, using \eqref{eq:sand_mu_bmu},
+which implies, using \eqref{eq:san_mu_bmu},
 \begin{align}
@@
   = S_\mu^\ell \p{S+\mu S_\z}\p{-\mu+S_\z}^m S_\bmu^n
-  - S_\mu^\ell\p{S + \f12\sp{\ell+n} + \mu S_\z} S_\z^m S_\bmu^n,
+  - S_\mu^\ell\sp{S + \f12\p{\ell+n} + \mu S_\z} S_\z^m S_\bmu^n,
   \label{eq:decay_diff}
 \end{align}
-and, using \eqref{eq:sand_bmu_mu},
+and, using \eqref{eq:san_bmu_mu},
 \begin{align}
@@
   &= S_\mu^\ell \p{S - \ell - n - \mu S_\z} \p{\mu + S_\z}^m S_\bmu^n
-  - S_\mu^\ell\p{S - \f12\sp{\ell+n} - \mu S_\z}
+  - S_\mu^\ell\sp{S - \f12\p{\ell+n} - \mu S_\z}
   S_\z^m S_\bmu^n \notag \\
@@
 
+\subsection{Dephasing}
+\label{sec:dephasing_single}
+
+The effect of decoherence via single-spin dephasing is described by
+\begin{align}
+  \D\p{\sigma_\z/\sqrt2; \O}
+  = \sum_j\p{\sigma_\z^{(j)} \O \sigma_\z^{(j)}
+    - \f12\sp{\sigma_\z^{(j)} \sigma_\z^{(j)},\O}_+}
+  = \f12 \sum_j \sigma_\z^{(j)} \O \sigma_\z^{(j)} - S \O,
+\end{align}
+where the factor $\sqrt2$ is inserted as
+e.g.~$\D\p{\sigma_\z/\sqrt2;S_\mu}=-S_\mu$, such that the above
+expression generates decay of single-weight transverse collective spin
+operators at a unit rate.  From \eqref{eq:san_z_z}, we then have
+\begin{align}
+  \D\p{\sigma_\z/\sqrt2; S_\mu^\ell S_\z^m S_\bmu^n}
+  = -\p{\ell+n} S_\mu^\ell S_\z^m S_\bmu^n
+  + 2 \ell n S_\mu^{\ell-1} \p{S + \mu S_\z}
+  \p{-\mu + S_\z}^m S_\bmu^{n-1}.
+\end{align}
+Decoherence via single-spin dephasing makes operators
+$S_\mu^\ell S_\z^m S_\bmu^n$ ``grow'' in $S_\z$, although the sum
+$\ell+m+n$ does not grow.
+
 \subsection{The general case}
@@
   = \abs{\gamma_\z}^2
-  + \sum_\mu \p{\abs{\gamma_\mu}^2 \f12 \sp{1-\mu\sigma_\z}
-    + \mu \sp{\gamma_\z^*\gamma_\mu - \gamma_\bmu^*\gamma_\z}
+  + \sum_\mu \sp{\abs{\gamma_\mu}^2 \f12 \p{1-\mu\sigma_\z}
+    + \mu \p{\gamma_\z^*\gamma_\mu - \gamma_\bmu^*\gamma_\z}
     \sigma_\mu},
@@
   &= \sum_{X\in\set{\z,+,-}} \abs{\gamma_X}^2 \D\p{\sigma_X;\O}
-  + \sum_{j,\mu}
+  + \sum_{\mu,j}
   \p{\gamma_\bmu^* \gamma_\mu \sigma_\mu^{(j)} \O \sigma_\mu^{(j)}
@@
   \sp{S_\mu, S_\mu^\ell S_\z^m S_\bmu^n}_+
-  &= S_\mu^{\ell+1} \sp{S_\z^m+\p{\mu+S_\z}^m} S_\bmu^n
+  = S_\mu^{\ell+1} \sp{S_\z^m+\p{\mu+S_\z}^m} S_\bmu^n
   - n S_\mu^\ell \p{n-1+2\mu S_\z} S_\z^m S_\bmu^{n-1}.
@@
 \end{align}
-This anti-commutator together with the results in Sections
-\ref{sec:sandwich_single}, \ref{sec:decay_single}, and
-\ref{sec:dephasing_single} are sufficient to expand
-\eqref{eq:D_general_single} in terms of ordered products of collective
-spin operators.
+Recognizing a resemblance between \eqref{eq:S_mu_acomm} and
+\eqref{eq:san_z_mu}, we can collect terms to simplify
+\begin{align}
+  \sum_j \sigma_\z^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\mu^{(j)}
+  - \f{\mu}{2} \sp{S_\mu, S_\mu^\ell S_\z^m S_\bmu^n}_+
+  = K_{\ell mn;\mu} + L_{\ell mn;\mu}
+  \label{eq:dec_z_mu}
+\end{align}
+and likewise
+\begin{align}
+  \sum_j \sigma_\mu^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\z^{(j)}
+  + \f{\mu}{2} \sp{S_\mu, S_\mu^\ell S_\z^m S_\bmu^n}_+
+  = K_{\ell mn;\mu} + M_{\ell mn;\mu}
+  \label{eq:dec_mu_z}
+\end{align}
+with
+\begin{align}
+  K_{\ell mn;\mu}
+  &\equiv \f{\mu}{2} S_\mu^{\ell+1}
+  \sp{\p{\mu+S_\z}^m-S_\z^m} S_\bmu^n \\
+  L_{\ell mn;\mu}
+  &\equiv \mu n S_\mu^\ell \sp{-2S+2\ell+\f32\p{n-1}+\mu S_\z}
+  S_\z^m S_\bmu^{n-1} \notag \\
+  &\qquad - 2\mu\ell n\p{n-1} S_\mu^{\ell-1}
+  \p{S+\mu S_\z} \p{-\mu+S_\z}^m S_\bmu^{n-2} \\
+  M_{\ell mn;\mu}
+  &\equiv \mu n S_\mu^\ell \sp{2\p{S+\mu S_\z}\p{-\mu+S_\z}^m
+    - \p{\f12\sp{n-1}+\mu S_\z} S_\z^m} S_\bmu^{n-1}.
+\end{align}
+Defining for completion
+\begin{align}
+  P_{\ell mn;\mu}
+  \equiv \sum_j
+  \sigma_\mu^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\mu^{(j)}
+  = n S_\mu^{\ell+1} S_\z^m S_\bmu^{n-1}
+  - n\p{n-1} S_\mu^\ell \p{S+\mu S_\z} \p{-\mu+S_\z}^m S_\bmu^{n-2},
+\end{align}
+and
+\begin{align}
+  Q_{\ell mn;\mu}^{(\gamma)}
+  \equiv \gamma_\bmu^* \gamma_\mu P_{\ell mn;\mu}
+  + \p{\gamma_\z^* \gamma_\mu + \gamma_\bmu^* \gamma_\z} K_{\ell mn;\mu}
+  + \gamma_\z^* \gamma_\mu L_{\ell mn;\mu}
+  + \gamma_\bmu^* \gamma_\z M_{\ell mn;\mu},
+\end{align}
+we finally have
+\begin{align}
+  \D\p{\gamma; S_\mu^\ell S_\z^m S_\bmu^n}
+  = \sum_{X\in\set{\z,+,-}} \abs{\gamma_X}^2
+  \D\p{\sigma_X; S_\mu^\ell S_\z^m S_\bmu^n}
+  + Q_{\ell mn;\mu}^{(\gamma)} + \sp{Q_{nm\ell;\mu}^{(\gamma)}}^\dag.
+\end{align}
 
@@
   S_\z S_\mu^\ell S_\z^m S_\bmu^n S_\z
-  &= S_\mu^\ell \p{\ell n + \mu\sp{\ell+n} S_\z + S_\z^2}
+  &= S_\mu^\ell \sp{\ell n + \mu\p{\ell+n} S_\z + S_\z^2}
   S_\z^m S_\bmu^n.
@@
   - n S_\mu^\ell \p{n-1+2\mu S_\z} S_\z^m S_\bmu^{n-1},
-  \allowdisplaybreaks \\
-  S_\bmu S_\mu^\ell S_\z^m S_\bmu^n
-  &= S_\mu^\ell \p{\mu+S_\z}^m S_\bmu^{n+1}
-  - \ell S_\mu^{\ell-1} \p{\ell-1+2\mu S_\z} S_\z^m S_\bmu^n,
 \end{align}
@@
   S_\z S_\mu^\ell S_\z^m S_\bmu^n S_\mu
-  &= S_\mu^{\ell+1} \p{\mu\sp{\ell+1}+S_\z} \p{\mu+S_\z}^m S_\bmu^n
+  &= S_\mu^{\ell+1} \sp{\mu\p{\ell+1}+S_\z} \p{\mu+S_\z}^m S_\bmu^n
   \notag \\
   &\qquad - n S_\mu^\ell
-  \p{\mu\ell\sp{n-1} + \sp{2\ell+n-1}S_\z + 2\mu S_\z^2}
+  \sp{\mu\ell\p{n-1} + \p{2\ell+n-1}S_\z + 2\mu S_\z^2}
   S_\z^m S_\bmu^{n-1}.
@@
   &= S_\mu^\ell S_\bmu S_\z^m S_\mu S_\bmu^n
-  - S_\mu^\ell \p{\ell\sp{\ell+1} + n\sp{n+1}+2\mu\sp{\ell+n} S_\z}
+  - S_\mu^\ell \sp{\ell\p{\ell+1} + n\p{n+1}+2\mu\p{\ell+n} S_\z}
   \p{\mu+S_\z}^m S_\bmu^n \notag \\
   &\qquad + \ell n S_\mu^{\ell-1}
-  \p{\sp{\ell-1}\sp{n-1}+2\mu\sp{\ell+n-2} S_\z + 4 S_\z^2}
+  \sp{\p{\ell-1}\p{n-1}+2\mu\p{\ell+n-2} S_\z + 4 S_\z^2}
   S_\z^m S_\bmu^{n-1},
@@
   S_\bmu S_\mu^\ell S_\z^m S_\bmu^n S_\mu
-  &= S_\mu^{\ell+1} \p{2\mu + S_\z}^m S_\bmu^{n+1}
-  - S_\mu^\ell \p{\ell\sp{\ell+1} + n\sp{n+1}+2\mu\sp{\ell+n+1} S_\z}
+  &= S_\mu^{\ell+1} \p{2\mu + S_\z}^m S_\bmu^{n+1} \notag \\
+  &\qquad - S_\mu^\ell
+  \sp{\ell\p{\ell+1} + n\p{n+1}+2\mu\p{\ell+n+1} S_\z}
   \p{\mu+S_\z}^m S_\bmu^n \notag \\
   &\qquad + \ell n S_\mu^{\ell-1}
-  \p{\sp{\ell-1}\sp{n-1}+2\mu\sp{\ell+n-2} S_\z + 4 S_\z^2}
+  \sp{\p{\ell-1}\p{n-1}+2\mu\p{\ell+n-2} S_\z + 4 S_\z^2}
   S_\z^m S_\bmu^{n-1}.
@@
   \D\p{S_\bmu; S_\mu^\ell S_\z^m S_\bmu^n}
-  &= -S_\mu^{\ell+1} \p{\sp{\mu+S_\z}^m - S_\z^m} S_\bmu^{n+1} \notag \\
-  &\qquad + \f12 S_\mu^\ell \p{\ell\sp{\ell-1} + n\sp{n-1}
-    + 2\mu\sp{\ell+n} S_\z} S_\z^m S_\bmu^n,
+  &= -S_\mu^{\ell+1} \sp{\p{\mu+S_\z}^m - S_\z^m} S_\bmu^{n+1} \notag \\
+  &\qquad + \f12 S_\mu^\ell \sp{\ell\p{\ell-1} + n\p{n-1}
+    + 2\mu\p{\ell+n} S_\z} S_\z^m S_\bmu^n,
 \end{align}
@@
   \D\p{S_\mu; S_\mu^\ell S_\z^m S_\bmu^n}
-  &= S_\mu^{\ell+1} \p{\sp{2\mu + S_\z}^m-\sp{\mu + S_\z}^m}
+  &= S_\mu^{\ell+1} \sp{\p{2\mu + S_\z}^m-\p{\mu + S_\z}^m}
   S_\bmu^{n+1} \notag \\
   &\qquad - S_\mu^\ell
-  \p{\ell\sp{\ell+1} + n\sp{n+1}+2\mu\sp{\ell+n+1} S_\z}
+  \sp{\ell\p{\ell+1} + n\p{n+1}+2\mu\p{\ell+n+1} S_\z}
   \p{\mu+S_\z}^m S_\bmu^n \notag \\
   &\qquad + \f12 S_\mu^\ell
-  \p{\ell\sp{\ell+1} + n\sp{n+1}+2\mu\sp{\ell+n+2} S_\z}
+  \sp{\ell\p{\ell+1} + n\p{n+1}+2\mu\p{\ell+n+2} S_\z}
   S_\z^m S_\bmu^n \notag \\
   &\qquad + \ell n S_\mu^{\ell-1}
-  \p{\sp{\ell-1}\sp{n-1}+2\mu\sp{\ell+n-2} S_\z + 4 S_\z^2}
+  \sp{\p{\ell-1}\p{n-1}+2\mu\p{\ell+n-2} S_\z + 4 S_\z^2}
   S_\z^m S_\bmu^{n-1}.
@@
   \notag \\
-  &\qquad -\f12 \sum_\mu\p{\Gamma_z^*\Gamma_\mu \sp{S_\z S_\mu, \O}_+
-    + \Gamma_\bmu^* \Gamma_z \sp{S_\mu S_\z, \O}_+
-    + \Gamma_\bmu^* \Gamma_\mu \sp{S_\mu^2, \O}_+}.
+  &\qquad -\f12 \sum_\mu\p{\Gamma_\bmu^* \Gamma_\mu \sp{S_\mu^2, \O}_+
+    + \Gamma_z^*\Gamma_\mu \sp{S_\z S_\mu, \O}_+
+    + \Gamma_\bmu^* \Gamma_z \sp{S_\mu S_\z, \O}_+}.
   \label{eq:D_general_collective}
@@
 \begin{align}
+  \sp{S_\mu^2, S_\mu^\ell S_\z^m S_\bmu^n}_+
+  &= S_\mu^{\ell+2} \sp{S_\z^m+\p{2\mu+S_\z}^m} S_\bmu^n
+  - 2n S_\mu^{\ell+1} \p{n+2\mu S_\z} \p{\mu+S_\z}^m S_\bmu^{n-1}
+  \notag \\
+  &\qquad + n\p{n-1} S_\mu^\ell \sp{\p{n-1}\p{n-2}
+    + 2\mu\p{2n-3} S_\z + 4 S_\z^2} S_\z^m S_\bmu^{n-2},
+  \allowdisplaybreaks \\
   \sp{S_\z S_\mu, S_\mu^\ell S_\z^m S_\bmu^n}_+
@@
     + \p{\mu\sp{n+1} + S_\z} \p{\mu+S_\z}^m } S_\bmu^n \notag \\
-  &\qquad - n S_\mu^\ell \p{\mu n \sp{n-1}
-    + \sp{3n-1} S_\z + 2\mu S_\z^2} S_\z^m S_\bmu^{n-1},
-  \allowdisplaybreaks \\
-  \sp{S_\bmu S_\z, S_\mu^\ell S_\z^m S_\bmu^n}_+
-  &= S_\mu^\ell \sp{\p{\mu\sp{\ell+1} + S_\z} \p{\mu+S_\z}^m
-    + \p{\mu\sp{n+1} + S_\z} S_\z^m } S_\bmu^{n+1} \notag \\
-  &\qquad - \ell S_\mu^{\ell-1} \p{\mu \ell \sp{\ell-1}
-    + \sp{3\ell-1} S_\z + 2\mu S_\z^2} S_\z^m S_\bmu^n,
+  &\qquad - n S_\mu^\ell \sp{\mu n \p{n-1}
+    + \p{3n-1} S_\z + 2\mu S_\z^2} S_\z^m S_\bmu^{n-1},
   \allowdisplaybreaks \\
@@
     + \p{\mu n+S_\z} \p{\mu+S_\z}^m} S_\bmu^n \notag \\
-  &\qquad - n S_\mu^\ell \p{\mu\sp{n-1}^2
-    + 3\sp{n-1} S_\z + 2\mu S_\z^2} S_\z^m S_\bmu^{n-1},
+  &\qquad - n S_\mu^\ell \sp{\mu\p{n-1}^2
+    + 3\p{n-1} S_\z + 2\mu S_\z^2} S_\z^m S_\bmu^{n-1}.
+\end{align}
+We can now define
+\begin{align}
+  \tilde P_{\ell mn;\mu}
+  &\equiv -\f12 S_\mu^{\ell+2}
+  \sp{\p{2\mu+S_\z}^m - 2\p{\mu+S_\z}^m + S_\z^m} S_\bmu^n \notag \\
+  &\qquad + n S_\mu^{\ell+1} \sp{\p{n+2\mu S_\z} \p{\mu+S_\z}^m
+    - \p{n-1+2\mu S_\z} S_\z^m} S_\bmu^{n-1} \notag \\
+  &\qquad -n\p{n-1} S_\mu^\ell
+  \sp{\f12\p{n-1}\p{n-2} + \mu\p{2n-3} S_\z + 2 S_\z^2}
+  S_\z^m S_\bmu^{n-2},
   \allowdisplaybreaks \\
-  \sp{S_\z S_\bmu, S_\mu^\ell S_\z^m S_\bmu^n}_+
-  &= S_\mu^\ell \sp{\p{\mu\ell+S_\z} \p{\mu+S_\z}^m
-    + \p{\mu n+S_\z} S_\z^m} S_\bmu^{n+1} \notag \\
-  &\qquad - \ell S_\mu^{\ell-1} \p{\mu\sp{\ell-1}^2
-    + 3\sp{\ell-1} S_\z + 2\mu S_\z^2} S_\z^m S_\bmu^n,
+  \tilde L_{\ell mn;\mu}^{(\Gamma)}
+  &\equiv \sp{\mu \p{2\ell-n+1} \Gamma_z^* \Gamma_\mu
+    - \mu n \Gamma_\bmu^*\Gamma_z}
+  S_\mu^{\ell+1} \p{\mu+S_\z}^m S_\bmu^n \notag \\
+  &\qquad + \p{\Gamma_z^* \Gamma_\mu - \Gamma_\bmu^*\Gamma_z}
+  S_\mu^{\ell+1} S_\z \p{\mu+S_\z}^m S_\bmu^n \notag \\
+  &\qquad - \sp{\mu\p{\ell+1} \Gamma_z^* \Gamma_\mu
+    - \mu\p{2n-\ell} \Gamma_\bmu^* \Gamma_z}
+  S_\mu^{\ell+1} S_\z^m S_\bmu^n \notag \\
+  &\qquad - \p{\Gamma_z^* \Gamma_\mu - \Gamma_\bmu^*\Gamma_z}
+  S_\mu^{\ell+1} S_\z^{m+1} S_\bmu^n,
   \allowdisplaybreaks \\
-  \sp{S_\mu^2, S_\mu^\ell S_\z^m S_\bmu^n}_+
-  &= S_\mu^{\ell+2} \p{S_\z^m+\sp{2\mu+S_\z}^m} S_\bmu^n
-  - 2n S_\mu^{\ell+1} \p{n+2\mu S_\z} \p{\mu+S_\z}^m S_\bmu^{n-1}
-  \notag \\
-  &\qquad + n\p{n-1} S_\mu^\ell \p{\sp{n-1}\sp{n-2}
-    + 2\mu\sp{2n-3} S_\z + 4 S_\z^2} S_\z^m S_\bmu^{n-2},
+  \tilde M_{\ell mn;\mu}^{(\Gamma)}
+  &= \sp{\mu\p{2\ell-n}\p{n-1} \Gamma_z^* \Gamma_\mu
+    - \mu\p{n-1}^2 \Gamma_\bmu^* \Gamma_z}
+  S_\mu^\ell S_\z^m S_\bmu^{n-1} \notag \\
+  &\qquad + \sp{\p{4\ell-n-1} \Gamma_z^* \Gamma_\mu
+    - 3\p{n-1} \Gamma_\bmu^* \Gamma_z}
+  S_\mu^\ell S_\z^{m+1} S_\bmu^{n-1} \notag \\
+  &\qquad +2\mu \p{\Gamma_z^* \Gamma_\mu - \Gamma_\bmu^* \Gamma_z}
+  S_\mu^\ell S_\z^{m+2} S_\bmu^{n-1},
   \allowdisplaybreaks \\
-  \sp{S_\bmu^2, S_\mu^\ell S_\z^m S_\bmu^n}_+
-  &= S_\mu^\ell \p{S_\z^m+\sp{2\mu+S_\z}^m} S_\bmu^{n+2}
-  - 2\ell S_\mu^{\ell-1} \p{\ell+2\mu S_\z} \p{\mu+S_\z}^m S_\bmu^{n+1}
-  \notag \\
-  &\qquad + \ell\p{\ell-1} S_\mu^{\ell-2} \p{\sp{\ell-1}\sp{\ell-2}
-    + 2\mu\sp{2\ell-3} S_\z + 4 S_\z^2} S_\z^m S_\bmu^n.
-\end{align}
-As with the case of general single-spin decoherence, there is not much
-to gain from substituting $\O = S_\mu^\ell S_\z^m S_\bmu^n$ into
-\eqref{eq:D_general_collective} and further simplifying the resulting
-expression; the anti-commutators provided in above together with the
-results in in Sections \ref{sec:sandwich_collective} and
-\ref{sec:decay_dephasing_collective} are sufficient to compute the
-effects of collective-spin decoherence when necessary.
+  \tilde Q_{\ell mn;\mu}^{(\Gamma)}
+  &\equiv \Gamma_\bmu^* \Gamma_\mu \tilde P_{\ell mn;\mu}
+  + \f12 \tilde L_{\ell mn;\mu}^{(\Gamma)}
+  - \f12 n \tilde M_{\ell mn;\mu}^{(\Gamma)},
+\end{align}
+in terms of which
+\begin{align}
+  \D\p{\Gamma; S_\mu^\ell S_\z^m S_\bmu^n}
+  = \sum_{X\in\set{\z,+,-}} \abs{\Gamma_X}^2
+  \D\p{S_X; S_\mu^\ell S_\z^m S_\bmu^n}
+  + \tilde Q_{\ell mn;\mu}^{(\Gamma)}
+  + \sp{\tilde Q_{nm\ell;\mu}^{(\Gamma)}}^\dag.
+\end{align}
 

### 2018-09-07 soc_squeezing/cumulant_expansions.tex
more simplifying of general decoherence calcualtions

@@
   &
-  \Gamma \equiv \gamma_z + \lambda,
+  \Gamma \equiv \gamma_\z + \lambda,
   &&
@@
 defined respectively in terms of spontaneous uncorrelated spin
-dephasing, excitation, and decay rates $\gamma_z$, $\gamma_+$,
+dephasing, excitation, and decay rates $\gamma_\z$, $\gamma_+$,
 $\gamma_-$.  The last ingredients we need to compute spin squeezing at
@@
   + \gamma_+ \sigma_+ + \gamma_- \sigma_-.
-  \label{eq:D_general_single_start}
-\end{align}
-To simplify \eqref{eq:D_general_single_start}, we expand
+  \label{eq:D_general_single}
+\end{align}
+To simplify \eqref{eq:D_general_single}, we expand
 \begin{align}
@@
   \p{\gamma_\z^*\gamma_\mu - \gamma_\bmu^*\gamma_\z} \sp{S_\mu, \O}_+.
-  \label{eq:D_general_single}
 \end{align}
@@
   - n S_\mu^\ell \p{n-1+2\mu S_\z} S_\z^m S_\bmu^{n-1}.
-  \label{eq:S_mu_acomm}
-\end{align}
-Recognizing a resemblance between \eqref{eq:S_mu_acomm} and
-\eqref{eq:san_z_mu}, we can collect terms to simplify
-\begin{align}
-  \sum_j \sigma_\z^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\mu^{(j)}
-  - \f{\mu}{2} \sp{S_\mu, S_\mu^\ell S_\z^m S_\bmu^n}_+
-  = K_{\ell mn;\mu} + L_{\ell mn;\mu}
-  \label{eq:dec_z_mu}
-\end{align}
-and likewise
-\begin{align}
-  \sum_j \sigma_\mu^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\z^{(j)}
-  + \f{\mu}{2} \sp{S_\mu, S_\mu^\ell S_\z^m S_\bmu^n}_+
-  = K_{\ell mn;\mu} + M_{\ell mn;\mu}
-  \label{eq:dec_mu_z}
-\end{align}
-with
-\begin{align}
+\end{align}
+Defining
+\begin{align}
+  \gamma_{\z,\mu}^{(\pm)}
+  &\equiv \f12\p{\gamma_\z^* \gamma_\mu \pm \gamma_\bmu^* \gamma_\z},
+  \allowdisplaybreaks \\
   K_{\ell mn;\mu}
-  &\equiv \f{\mu}{2} S_\mu^{\ell+1}
-  \sp{\p{\mu+S_\z}^m-S_\z^m} S_\bmu^n \\
-  L_{\ell mn;\mu}
-  &\equiv \mu n S_\mu^\ell \sp{-2S+2\ell+\f32\p{n-1}+\mu S_\z}
-  S_\z^m S_\bmu^{n-1} \notag \\
-  &\qquad - 2\mu\ell n\p{n-1} S_\mu^{\ell-1}
-  \p{S+\mu S_\z} \p{-\mu+S_\z}^m S_\bmu^{n-2} \\
-  M_{\ell mn;\mu}
-  &\equiv \mu n S_\mu^\ell \sp{2\p{S+\mu S_\z}\p{-\mu+S_\z}^m
-    - \p{\f12\sp{n-1}+\mu S_\z} S_\z^m} S_\bmu^{n-1}.
-\end{align}
-Defining for completion
-\begin{align}
+  &\equiv S_\mu^\ell \p{S+\mu S_\z} \p{-\mu+S_\z}^m S_\bmu^n,
+  \allowdisplaybreaks \\
+  L_{\ell mn;\mu}^{(\gamma)}
+  &\equiv \gamma_{\z,\mu}^{(+)} S_\mu^{\ell+1}
+  \sp{\p{\mu+S_\z}^m-S_\z^m} S_\bmu^n
+  + 2n \gamma_\bmu^* \gamma_z K_{\ell m,n-1;\mu}
+  - 2\ell n\p{n-1} \gamma_z^* \gamma_\mu K_{\ell-1,m,n-2;\mu},
+  \allowdisplaybreaks \\
+  M_{\ell mn;\mu}^{(\gamma)}
+  &\equiv \gamma_{\z,\mu}^{(-)} S_\mu^\ell
+  S_\z^{m+1} S_\bmu^{n-1}
+  -\mu \sp{\p{S-\ell-\f12n+\f12} \gamma_\z^* \gamma_\mu
+    - \f12 \p{n-1} \gamma_{\z,\mu}^{(-)}}
+  S_\mu^\ell S_\z^m S_\bmu^{n-1},
+  \allowdisplaybreaks \\
   P_{\ell mn;\mu}
-  \equiv \sum_j
-  \sigma_\mu^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\mu^{(j)}
-  = n S_\mu^{\ell+1} S_\z^m S_\bmu^{n-1}
-  - n\p{n-1} S_\mu^\ell \p{S+\mu S_\z} \p{-\mu+S_\z}^m S_\bmu^{n-2},
-\end{align}
-and
-\begin{align}
+  &\equiv S_\mu^{\ell+1} S_\z^m S_\bmu^{n-1}
+  - \p{n-1} K_{\ell m,n-2;\mu},
+  \allowdisplaybreaks \\
   Q_{\ell mn;\mu}^{(\gamma)}
-  \equiv \gamma_\bmu^* \gamma_\mu P_{\ell mn;\mu}
-  + \p{\gamma_\z^* \gamma_\mu + \gamma_\bmu^* \gamma_\z} K_{\ell mn;\mu}
-  + \gamma_\z^* \gamma_\mu L_{\ell mn;\mu}
-  + \gamma_\bmu^* \gamma_\z M_{\ell mn;\mu},
-\end{align}
-we finally have
+  &\equiv \mu L_{\ell mn;\mu}^{(\gamma)}
+  + 2 n M_{\ell mn;\mu}^{(\gamma)}
+  + n \gamma_\bmu^* \gamma_\mu P_{\ell mn;\mu},
+\end{align}
+we then have
 \begin{align}
@@
   + \sum_\mu \p{\abs{\Gamma_\mu}^2 S_\bmu S_\mu
-    + \Gamma_z^*\Gamma_\mu S_\z S_\mu
-    + \Gamma_\bmu^* \Gamma_z S_\mu S_\z
+    + \Gamma_\z^*\Gamma_\mu S_\z S_\mu
+    + \Gamma_\bmu^* \Gamma_\z S_\mu S_\z
     + \Gamma_\bmu^* \Gamma_\mu S_\mu^2},
@@
   &\qquad -\f12 \sum_\mu\p{\Gamma_\bmu^* \Gamma_\mu \sp{S_\mu^2, \O}_+
-    + \Gamma_z^*\Gamma_\mu \sp{S_\z S_\mu, \O}_+
-    + \Gamma_\bmu^* \Gamma_z \sp{S_\mu S_\z, \O}_+}.
-  \label{eq:D_general_collective}
+    + \Gamma_\z^*\Gamma_\mu \sp{S_\z S_\mu, \O}_+
+    + \Gamma_\bmu^* \Gamma_\z \sp{S_\mu S_\z, \O}_+}.
 \end{align}
@@
 \end{align}
-We can now define
-\begin{align}
+Defining
+\begin{align}
+  \Gamma_{\z,\mu}^{(\pm)}
+  &\equiv \f12\p{\Gamma_\z^* \Gamma_\mu \pm \Gamma_\bmu^* \Gamma_\z},
+  \allowdisplaybreaks \\
+  \tilde L_{\ell mn;\mu}^{(\Gamma)}
+  &\equiv \mu \sp{\p{\ell-n+\f12} \Gamma_{\z,\mu}^{(+)}
+    + \p{\ell+\f12} \Gamma_{\z,\mu}^{(-)}}
+  S_\mu^{\ell+1} \p{\mu+S_\z}^m S_\bmu^n \notag \\
+  &\qquad -\mu \sp{\p{\ell-n+\f12} \Gamma_{\z,\mu}^{(+)}
+    + \p{n+\f12} \Gamma_{\z,\mu}^{(-)}}
+  S_\mu^{\ell+1} S_\z^m S_\bmu^n \notag \\
+  &\qquad + \Gamma_{\z,\mu}^{(-)}
+  S_\mu^{\ell+1} S_\z \sp{\p{\mu+S_\z}^m - S_\z^m} S_\bmu^n,
+  \allowdisplaybreaks \\
+  \tilde M_{\ell mn;\mu}^{(\Gamma)}
+  &= \mu \p{n-1} \sp{\p{\ell-n+\f12} \Gamma_{\z,\mu}^{(+)}
+    + \p{\ell-\f12} \Gamma_{\z,\mu}^{(-)}}
+  S_\mu^\ell S_\z^m S_\bmu^{n-1} \notag \\
+  &\qquad + 2 \sp{\p{\ell-n+\f12} \Gamma_{\z,\mu}^{(+)}
+    + \p{\ell+\f12n-1} \Gamma_{\z,\mu}^{(-)}}
+  S_\mu^\ell S_\z^{m+1} S_\bmu^{n-1} \notag \\
+  &\qquad + 2\mu \Gamma_{\z,\mu}^{(-)} S_\mu^\ell S_\z^{m+2} S_\bmu^{n-1},
+  \allowdisplaybreaks \\
   \tilde P_{\ell mn;\mu}
@@
   \allowdisplaybreaks \\
-  \tilde L_{\ell mn;\mu}^{(\Gamma)}
-  &\equiv \sp{\mu \p{2\ell-n+1} \Gamma_z^* \Gamma_\mu
-    - \mu n \Gamma_\bmu^*\Gamma_z}
-  S_\mu^{\ell+1} \p{\mu+S_\z}^m S_\bmu^n \notag \\
-  &\qquad + \p{\Gamma_z^* \Gamma_\mu - \Gamma_\bmu^*\Gamma_z}
-  S_\mu^{\ell+1} S_\z \p{\mu+S_\z}^m S_\bmu^n \notag \\
-  &\qquad - \sp{\mu\p{\ell+1} \Gamma_z^* \Gamma_\mu
-    - \mu\p{2n-\ell} \Gamma_\bmu^* \Gamma_z}
-  S_\mu^{\ell+1} S_\z^m S_\bmu^n \notag \\
-  &\qquad - \p{\Gamma_z^* \Gamma_\mu - \Gamma_\bmu^*\Gamma_z}
-  S_\mu^{\ell+1} S_\z^{m+1} S_\bmu^n,
-  \allowdisplaybreaks \\
-  \tilde M_{\ell mn;\mu}^{(\Gamma)}
-  &= \sp{\mu\p{2\ell-n}\p{n-1} \Gamma_z^* \Gamma_\mu
-    - \mu\p{n-1}^2 \Gamma_\bmu^* \Gamma_z}
-  S_\mu^\ell S_\z^m S_\bmu^{n-1} \notag \\
-  &\qquad + \sp{\p{4\ell-n-1} \Gamma_z^* \Gamma_\mu
-    - 3\p{n-1} \Gamma_\bmu^* \Gamma_z}
-  S_\mu^\ell S_\z^{m+1} S_\bmu^{n-1} \notag \\
-  &\qquad +2\mu \p{\Gamma_z^* \Gamma_\mu - \Gamma_\bmu^* \Gamma_z}
-  S_\mu^\ell S_\z^{m+2} S_\bmu^{n-1},
-  \allowdisplaybreaks \\
   \tilde Q_{\ell mn;\mu}^{(\Gamma)}
-  &\equiv \Gamma_\bmu^* \Gamma_\mu \tilde P_{\ell mn;\mu}
-  + \f12 \tilde L_{\ell mn;\mu}^{(\Gamma)}
-  - \f12 n \tilde M_{\ell mn;\mu}^{(\Gamma)},
-\end{align}
-in terms of which
+  &\equiv \tilde L_{\ell mn;\mu}^{(\Gamma)}
+  - n \tilde M_{\ell mn;\mu}^{(\Gamma)}
+  + \Gamma_\bmu^* \Gamma_\mu \tilde P_{\ell mn;\mu},
+\end{align}
+we then have
 \begin{align}

### 2018-09-07 soc_squeezing/cumulant_expansions.tex
added alternate method for computing arbitrary single-spin decoherence

@@
   - n S_\mu^\ell \p{n-1+2\mu S_\z} S_\z^m S_\bmu^{n-1}.
-\end{align}
-Defining
-\begin{align}
-  \gamma_{\z,\mu}^{(\pm)}
-  &\equiv \f12\p{\gamma_\z^* \gamma_\mu \pm \gamma_\bmu^* \gamma_\z},
-  \allowdisplaybreaks \\
+  \label{eq:S_mu_acomm}
+\end{align}
+Recognizing a resemblance between \eqref{eq:S_mu_acomm} and
+\eqref{eq:san_z_mu}, we can collect terms to simplify
+\begin{align}
+  \sum_j \sigma_\z^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\mu^{(j)}
+  - \f{\mu}{2} \sp{S_\mu, S_\mu^\ell S_\z^m S_\bmu^n}_+
+  = K_{\ell mn;\mu} + L_{\ell mn;\mu}
+  \label{eq:dec_z_mu}
+\end{align}
+and likewise
+\begin{align}
+  \sum_j \sigma_\mu^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\z^{(j)}
+  + \f{\mu}{2} \sp{S_\mu, S_\mu^\ell S_\z^m S_\bmu^n}_+
+  = K_{\ell mn;\mu} + M_{\ell mn;\mu}
+  \label{eq:dec_mu_z}
+\end{align}
+with
+\begin{align}
   K_{\ell mn;\mu}
-  &\equiv S_\mu^\ell \p{S+\mu S_\z} \p{-\mu+S_\z}^m S_\bmu^n,
-  \allowdisplaybreaks \\
-  L_{\ell mn;\mu}^{(\gamma)}
-  &\equiv \gamma_{\z,\mu}^{(+)} S_\mu^{\ell+1}
-  \sp{\p{\mu+S_\z}^m-S_\z^m} S_\bmu^n
-  + 2n \gamma_\bmu^* \gamma_z K_{\ell m,n-1;\mu}
-  - 2\ell n\p{n-1} \gamma_z^* \gamma_\mu K_{\ell-1,m,n-2;\mu},
-  \allowdisplaybreaks \\
-  M_{\ell mn;\mu}^{(\gamma)}
-  &\equiv \gamma_{\z,\mu}^{(-)} S_\mu^\ell
-  S_\z^{m+1} S_\bmu^{n-1}
-  -\mu \sp{\p{S-\ell-\f12n+\f12} \gamma_\z^* \gamma_\mu
-    - \f12 \p{n-1} \gamma_{\z,\mu}^{(-)}}
-  S_\mu^\ell S_\z^m S_\bmu^{n-1},
-  \allowdisplaybreaks \\
+  &\equiv \f{\mu}{2} S_\mu^{\ell+1}
+  \sp{\p{\mu+S_\z}^m-S_\z^m} S_\bmu^n, \\
+  L_{\ell mn;\mu}
+  &\equiv \mu n S_\mu^\ell \sp{-2S+2\ell+\f32\p{n-1}+\mu S_\z}
+  S_\z^m S_\bmu^{n-1} \notag \\
+  &\qquad - 2\mu\ell n\p{n-1} S_\mu^{\ell-1}
+  \p{S+\mu S_\z} \p{-\mu+S_\z}^m S_\bmu^{n-2}, \\
+  M_{\ell mn;\mu}
+  &\equiv \mu n S_\mu^\ell \sp{2\p{S+\mu S_\z}\p{-\mu+S_\z}^m
+    - \p{\f12\sp{n-1}+\mu S_\z} S_\z^m} S_\bmu^{n-1}.
+\end{align}
+Defining for completion
+\begin{align}
   P_{\ell mn;\mu}
-  &\equiv S_\mu^{\ell+1} S_\z^m S_\bmu^{n-1}
-  - \p{n-1} K_{\ell m,n-2;\mu},
-  \allowdisplaybreaks \\
+  \equiv \sum_j
+  \sigma_\mu^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\mu^{(j)}
+  = n S_\mu^{\ell+1} S_\z^m S_\bmu^{n-1}
+  - n\p{n-1} S_\mu^\ell \p{S+\mu S_\z} \p{-\mu+S_\z}^m S_\bmu^{n-2},
+\end{align}
+and
+\begin{align}
   Q_{\ell mn;\mu}^{(\gamma)}
-  &\equiv \mu L_{\ell mn;\mu}^{(\gamma)}
-  + 2 n M_{\ell mn;\mu}^{(\gamma)}
-  + n \gamma_\bmu^* \gamma_\mu P_{\ell mn;\mu},
-\end{align}
-we then have
+  \equiv \gamma_\bmu^* \gamma_\mu P_{\ell mn;\mu}
+  + \p{\gamma_\z^* \gamma_\mu + \gamma_\bmu^* \gamma_\z}
+  K_{\ell mn;\mu}
+  + \gamma_\z^* \gamma_\mu L_{\ell mn;\mu}
+  + \gamma_\bmu^* \gamma_\z M_{\ell mn;\mu},
+  \label{eq:Q_single}
+\end{align}
+we finally have
 \begin{align}
@@
   + Q_{\ell mn;\mu}^{(\gamma)} + \sp{Q_{nm\ell;\mu}^{(\gamma)}}^\dag.
+\end{align}
+Alternately, we can define
+\begin{align}
+  \gamma_{\z,\mu}^{(\pm)}
+  &\equiv \f12\p{\gamma_\z^* \gamma_\mu \pm \gamma_\bmu^* \gamma_\z},
+  \allowdisplaybreaks \\
+  R_{\ell mn;\mu}
+  &\equiv S_\mu^\ell \p{S+\mu S_\z} \p{-\mu+S_\z}^m S_\bmu^n,
+  \allowdisplaybreaks \\
+  L_{\ell mn;\mu}^{(\gamma)}
+  &\equiv \mu \gamma_{\z,\mu}^{(+)} S_\mu^{\ell+1}
+  \sp{\p{\mu+S_\z}^m-S_\z^m} S_\bmu^n
+  + 2\mu n \gamma_\bmu^* \gamma_z R_{\ell m,n-1;\mu} \notag \\
+  &\qquad - 2\mu \ell n\p{n-1} \gamma_z^* \gamma_\mu
+  R_{\ell-1,m,n-2;\mu},
+  \allowdisplaybreaks \\
+  M_{\ell mn;\mu}^{(\gamma)}
+  &\equiv 2n \gamma_{\z,\mu}^{(-)} S_\mu^\ell S_\z^{m+1} S_\bmu^{n-1}
+  \notag \\
+  &\qquad - 2\mu n \sp{\p{S-\ell-\f12n+\f12} \gamma_\z^* \gamma_\mu
+    - \f12 \p{n-1} \gamma_{\z,\mu}^{(-)}}
+  S_\mu^\ell S_\z^m S_\bmu^{n-1},
+\end{align}
+in terms of which
+\begin{align}
+  Q_{\ell mn;\mu}^{(\gamma)}
+  &\equiv L_{\ell mn;\mu}^{(\gamma)} + M_{\ell mn;\mu}^{(\gamma)}
+  + \gamma_\bmu^* \gamma_\mu P_{\ell mn;\mu}.
 \end{align}

### 2018-09-07 soc_squeezing/scripts/test_decoherence.py
removed obsolete method for single-spin decoherence calculations

@@ -138,43 +138,10 @@ for l, m, n in itertools.product(range(N), repeat = 3):
                      + gg_L * L(l,m,n,mu)
                      + gg_M * M(l,m,n,mu) )
 
-        def KK(l,m,n,mu):
-            return S_mu**l * (S+mu*S_z) * (-mu+S_z)**m * S_nu**n
-
-        def LL(l,m,n,mu):
-            L = gg_p * S_mu**(l+1) * ((mu+S_z)**m - S_z**m) * S_nu**n
-            if n >= 1:
-                L += 2*n * g_mz * KK(l,m,n-1,mu)
-            if n >= 2 and l >= 1:
-                L += -2*l*n*(n-1) * g_zp * KK(l-1,m,n-2,mu)
-            return L
-
-        def MM(l,m,n,mu):
-            if n == 0: return 0
-            M = gg_m * S_mu**l * S_z**(m+1) * S_nu**(n-1)
-            fac = (S-l-n/2+1/2) * g_zp - 1/2*(n-1) * gg_m
-            M += -mu * fac * S_mu**l * S_z**m * S_nu**(n-1)
-            return M
-
-        def PP(l,m,n,mu):
-            if n == 0: return 0
-            P = S_mu**(l+1) * S_z**m * S_nu**(n-1)
-            if n >= 2:
-                P += -(n-1) * KK(l,m,n-2,mu)
-            return P
-
-        def QQ(l,m,n,mu):
-            return mu * LL(l,m,n,mu) + 2*n * MM(l,m,n,mu) + n * g_mp * PP(l,m,n,mu)
-
         ##############################
 
         b = abs(g_z)**2 * d_z + abs(g_p)**2 * d_p + abs(g_m)**2 * d_m
         b += Q(l,m,n,mu) + Q(n,m,l,mu).dag()
-
-        check(a == b)
-
-        b = abs(g_z)**2 * d_z + abs(g_p)**2 * d_p + abs(g_m)**2 * d_m
-        b += QQ(l,m,n,mu) + QQ(n,m,l,mu).dag()
 
         check(a == b)
 

### 2018-09-08 soc_squeezing/scripts/correlator_methods.py
implemented and tested arbitrary single-weight collective-spin decoherence

@@ -10,11 +10,13 @@ from mpmath import hyper
 from scipy.special import factorial, binom
 from sympy.functions.combinatorial.numbers import stirling as sympy_stirling
 
+
+##########################################################################################
+# initial conditions
+##########################################################################################
+
 # natural logarithm of factorial
 def ln_factorial(n): return lgamma(n+1)
-
-# unsigned stirling number of the first kind
-def stirling(n,k): return float(sympy_stirling(n, k, kind = 1, signed = True))
 
 # correlator < +X | S_\mu^ll S_\z^mm S_\bmu^nn | +X >
 def op_val_pX(total_spin, op, mu):
@@ -44,6 +46,14 @@ def op_val_nZ(total_spin, op, mu = None):
     if ll != 0 or nn != 0: return 0
     return (-total_spin)**mm
 
+
+##########################################################################################
+# general commutator between ordered products of collective spin operators
+##########################################################################################
+
+# unsigned stirling number of the first kind
+def stirling(n,k): return float(sympy_stirling(n, k, kind = 1, signed = True))
+
 # collective spin operator commutator coefficients (see notes)
 def epsilon(mm,nn,pp,ll):
     return 2**ll * sum([ stirling(pp,qq) * binom(qq,ll) * (mm-nn)**(qq-ll)
@@ -72,17 +82,30 @@ def simplify_term(op_left, op_right, mu, prefactor = 1):
                     term[op_in] += klmn_sign * klmn_fac * prefactor
     return term
 
+
+##########################################################################################
+# machinery for decoherence codes
+##########################################################################################
+
 # take hermitian conjugate of a dictionary taking operator --> value,
[…]

### 2018-09-08 soc_squeezing/scripts/test_decoherence.py

@@ -170,12 +170,11 @@ for l, m, n in itertools.product(range(N), repeat = 3):
 
         def M(l,m,n,mu):
             if n == 0: return 0
-            M = ( mu * (n-1) * ( (l-n+1/2) * gg_p + (l-1/2) * gg_m )
+            M = ( -mu*n*(n-1) * ( (l-n+1/2) * gg_p + (l-1/2) * gg_m )
                   * S_mu**l * S_z**m * S_nu**(n-1) )
-            M += ( 2 * ( (l-n+1/2) * gg_p + (l+n/2-1) * gg_m )
+            M += ( -2*n * ( (l-n+1/2) * gg_p + (l+n/2-1) * gg_m )
                    * S_mu**l * S_z**(m+1) * S_nu**(n-1) )
-
-            M += 2 * mu * gg_m * S_mu**l * S_z**(m+2) * S_nu**(n-1)
+            M += -2*mu*n * gg_m * S_mu**l * S_z**(m+2) * S_nu**(n-1)
             return M
 
         def P(l,m,n,mu):
@@ -191,7 +190,7 @@ for l, m, n in itertools.product(range(N), repeat = 3):
             return P
 
         def Q(l,m,n,mu):
-            return L(l,m,n,mu) - n * M(l,m,n,mu) + np.conj(g_m) * g_p * P(l,m,n,mu)
+            return L(l,m,n,mu) + M(l,m,n,mu) + np.conj(g_m) * g_p * P(l,m,n,mu)
 
         ##############################
 

### 2018-09-08 soc_squeezing/cumulant_expansions.tex
removed obsolete method for single-spin decoherence calculations
implemented and tested arbitrary single-weight collective-spin decoherence
relabeled some indices in general commutator

--- 2018-09-07 soc_squeezing/cumulant_expansions.tex
@@
 \begin{align}
-  H = \sum_{p,q,r} h_{pqr} S_+^p S_\z^q S_-^r,
+  H = \sum_{\ell,m,n} h_{\ell mn} S_+^\ell S_\z^m S_-^n,
   \label{eq:H_general}
 \end{align}
-where Hermiticity requires that $h_{pqr}=h_{rqp}$.  The Hamiltonian
-for one-axis twisting with a transverse field, for example, is
+where Hermiticity requires that $h_{\ell mn}=h_{nm\ell}$.  The
+Hamiltonian for one-axis twisting with a transverse field, for
+example, is
 \begin{align}
@@
 \begin{align}
-  \sp{S_\mu^p S_\z^a S_\bmu^q, S_\mu^r S_\z^b S_\bmu^s}_\pm
-  = \S^{paq}_{rbs;\mu} \pm \S^{rbs}_{paq;\mu},
+  \sp{S_\mu^p S_\z^q S_\bmu^r, S_\mu^\ell S_\z^m S_\bmu^n}_\pm
+  = \S^{pqr}_{\ell mn;\mu} \pm \S^{\ell mn}_{pqr;\mu},
 \end{align}
@@
 \begin{align}
-  \S^{paq}_{rbs;\mu}
-  \equiv S_\mu^p S_\z^a S_\bmu^q S_\mu^r S_\z^b S_\bmu^s
-  = \sum_{k=0}^{\min\set{q,r}} k! { q \choose k } { r \choose k }
-  S_\mu^{p+r-k} \tilde Z_{qrab;\mu}^{(k)} S_\bmu^{q+s-k},
-\end{align}
-and, in terms of $\xi_{qrk}^{(\ell)}$ as defined in \eqref{eq:xi_mnp},
-\begin{align}
-  \tilde Z_{qrab;\mu}^{(k)}
-  &\equiv \p{\mu\sp{r-k}+S_\z}^a
-  Z_{qr;\bmu}^{(k)} \p{\mu\sp{q-k}+S_\z}^b \\
-  &= \sum_{\ell=0}^k \p{-1}^\ell \xi_{qrk}^{(\ell)}
-  \sum_{m=0}^a \p{r-k}^{a-m} { a \choose m }
-  \sum_{n=0}^b \p{q-k}^{b-n} { b \choose n }
-  \mu^{a+b+\ell-m-n} S_\z^{\ell+m+n}.
+  \S^{pqr}_{\ell mn;\mu}
+  \equiv S_\mu^p S_\z^q S_\bmu^r S_\mu^\ell S_\z^m S_\bmu^n
+  = \sum_{k=0}^{\min\set{r,\ell}} k! { r \choose k } { \ell \choose k }
+  S_\mu^{p+\ell-k} \tilde Z_{qr\ell m;\mu}^{(k)} S_\bmu^{r+n-k},
+\end{align}
+and, in terms of $\xi_{r\ell k}^{(a)}$ as defined in
+\eqref{eq:xi_mnp},
+\begin{align}
+  \tilde Z_{qr\ell m;\mu}^{(k)}
+  &\equiv \p{\mu\sp{\ell-k}+S_\z}^q
+  Z_{r\ell;\bmu}^{(k)} \p{\mu\sp{r-k}+S_\z}^m \\
+  &= \sum_{a=0}^k \p{-1}^a \xi_{r\ell k}^{(a)}
+  \sum_{b=0}^q \p{\ell-k}^{q-b} { q \choose b }
+  \sum_{c=0}^m \p{r-k}^{m-c} { m \choose c }
+  \mu^{q+m+a-b-c} S_\z^{a+b+c}.
 \end{align}
@@
 We are now equipped to derive similar identities for general ordered
-products of collective spin operators.  Making heavy use of identities
-\eqref{eq:push_z_mu_sS} and \eqref{eq:push_mu_Ss} to push single-spin
-operators through transverse collective-spin operators, we work
-through all combinations that are unique up to Hermitian conjugation,
-finding
+products of collective spin operators.  Making (heavy) use of
+identities \eqref{eq:push_z_mu_sS} and \eqref{eq:push_mu_Ss} to push
+single-spin operators through transverse collective-spin operators, we
+work through all combinations that are unique up to Hermitian
+conjugation, finding
 \begin{align}
@@
   \D\p{\sigma_\z/\sqrt2; \O}
-  = \sum_j\p{\sigma_\z^{(j)} \O \sigma_\z^{(j)}
+  = \f12 \sum_j\p{\sigma_\z^{(j)} \O \sigma_\z^{(j)}
     - \f12\sp{\sigma_\z^{(j)} \sigma_\z^{(j)},\O}_+}
@@
 \end{align}
-Defining
+Collecting terms and defining
 \begin{align}
@@
   \tilde M_{\ell mn;\mu}^{(\Gamma)}
-  &= \mu \p{n-1} \sp{\p{\ell-n+\f12} \Gamma_{\z,\mu}^{(+)}
+  &= -\mu n\p{n-1} \sp{\p{\ell-n+\f12} \Gamma_{\z,\mu}^{(+)}
     + \p{\ell-\f12} \Gamma_{\z,\mu}^{(-)}}
   S_\mu^\ell S_\z^m S_\bmu^{n-1} \notag \\
-  &\qquad + 2 \sp{\p{\ell-n+\f12} \Gamma_{\z,\mu}^{(+)}
+  &\qquad - 2n \sp{\p{\ell-n+\f12} \Gamma_{\z,\mu}^{(+)}
     + \p{\ell+\f12n-1} \Gamma_{\z,\mu}^{(-)}}
   S_\mu^\ell S_\z^{m+1} S_\bmu^{n-1} \notag \\
-  &\qquad + 2\mu \Gamma_{\z,\mu}^{(-)} S_\mu^\ell S_\z^{m+2} S_\bmu^{n-1},
+  &\qquad - 2\mu n \Gamma_{\z,\mu}^{(-)}
+  S_\mu^\ell S_\z^{m+2} S_\bmu^{n-1},
   \allowdisplaybreaks \\
@@
   \tilde Q_{\ell mn;\mu}^{(\Gamma)}
-  &\equiv \tilde L_{\ell mn;\mu}^{(\Gamma)}
-  - n \tilde M_{\ell mn;\mu}^{(\Gamma)}
-  + \Gamma_\bmu^* \Gamma_\mu \tilde P_{\ell mn;\mu},
+  &\equiv \Gamma_\bmu^* \Gamma_\mu \tilde P_{\ell mn;\mu}
+  + \tilde L_{\ell mn;\mu}^{(\Gamma)}
+  + \tilde M_{\ell mn;\mu}^{(\Gamma)},
 \end{align}

### 2018-09-08 soc_squeezing/scripts/correlator_methods.py

@@ -64,22 +64,22 @@ def xi(mm,nn,pp,qq):
 
 # simplify product of two operators
 def simplify_term(op_left, op_right, mu, prefactor = 1):
-    pp, aa, qq = op_left
-    rr, bb, ss = op_right
+    pp, qq, rr = op_left
+    ll, mm, nn = op_right
     term = {}
-    for kk in range(min(qq,rr)+1):
-        kk_fac = factorial(kk) * binom(qq,kk) * binom(rr,kk)
-        for ll in range(kk+1):
-            kl_fac = kk_fac * (-1)**ll * xi(qq,rr,kk,ll)
-            for mm in range(aa+1):
-                klm_fac = kl_fac * (rr-kk)**(aa-mm) * binom(aa,mm)
-                for nn in range(bb+1):
-                    klmn_fac = klm_fac * (qq-kk)**(bb-nn) * binom(bb,nn)
-                    klmn_sign = mu**(aa+bb+ll-mm-nn)
-                    op_in = (pp+rr-kk, ll+mm+nn, qq+ss-kk)
+    for kk in range(min(rr,ll)+1):
+        kk_fac = factorial(kk) * binom(rr,kk) * binom(ll,kk)
+        for aa in range(kk+1):
+            ka_fac = kk_fac * (-1)**aa * xi(rr,ll,kk,aa)
+            for bb in range(qq+1):
+                kab_fac = ka_fac * (ll-kk)**(qq-bb) * binom(qq,bb)
+                for cc in range(mm+1):
+                    kabc_fac = kab_fac * (rr-kk)**(mm-cc) * binom(mm,cc)
+                    kabc_sign = mu**(qq+mm+aa-bb-cc)
+                    op_in = (pp+ll-kk, aa+bb+cc, rr+nn-kk)
                     try: term[op_in]
                     except: term[op_in] = 0
-                    term[op_in] += klmn_sign * klmn_fac * prefactor
+                    term[op_in] += kabc_sign * kabc_fac * prefactor
     return term
 
 

### 2018-09-11 soc_squeezing/cumulant_expansions.tex
removed obsolete text for expectation value calculations, simplified expectation value for X added expectation value for any state in the X-Y plane
simplified expression for xi in commutator, and renamed xi --> zeta fixed incorrect expectation value for -Z state with mu = -1 recycle values of ln_factorial which we have already computed

@@
 \newcommand{\X}{\text{X}}
+\newcommand{\Y}{\text{Y}}
 \newcommand{\bmu}{{\bar\mu}}
+\newcommand{\bnu}{{\bar\nu}}
 \newcommand{\floor}[1]{\left\lfloor{#1}\right\rfloor}
@@
 Examining the nonzero entries of $T^{\ell mn}_{pqr;k}$ for
-$k=1,2,\cdots,M-1$ now {\it tells} us which operators we need to keep
+$k=0,1,\cdots,M-1$ now {\it tells} us which operators we need to keep
 track of in order to compute the expectation value
@@
   \p{-\mu\sp{m-p} + S_\z}^\ell
-  = \sum_{q=0}^p \mu^q \xi_{mnp}^{(q)} S_\z^q,
+  = \sum_{q=0}^p \mu^q \zeta_{mnp}^{(q)} S_\z^q,
   \label{eq:Z_mn}
@@
 \begin{align}
-  \xi_{mnp}^{(q)}
+  \zeta_{mnp}^{(q)}
   \equiv \sum_{\ell=q}^p \p{-1}^{\ell-q}
   \epsilon_{mnp}^{(\ell)} { \ell \choose q } \p{m-p}^{\ell-q}
-  \label{eq:xi_mnp}
-\end{align}
-in terms of $\epsilon_{mnp}^{(\ell)}$ as defined in
-\eqref{eq:eps_mnp}.
+  = \p{-1}^{p-q} 2^q \sum_{s=q}^p
+  { p \brack s } { s \choose q } \p{m+n-2p}^{s-q}.
+  \label{eq:zeta_mnp}
+\end{align}
 
@@
   S_\mu^{p+\ell-k} \tilde Z_{qr\ell m;\mu}^{(k)} S_\bmu^{r+n-k},
-\end{align}
-and, in terms of $\xi_{r\ell k}^{(a)}$ as defined in
-\eqref{eq:xi_mnp},
+  \label{eq:general_product}
+\end{align}
+and, in terms of $\zeta_{r\ell k}^{(a)}$ as defined in
+\eqref{eq:zeta_mnp},
 \begin{align}
@@
   Z_{r\ell;\bmu}^{(k)} \p{\mu\sp{r-k}+S_\z}^m \\
-  &= \sum_{a=0}^k \p{-1}^a \xi_{r\ell k}^{(a)}
+  &= \sum_{a=0}^k \p{-1}^a \zeta_{r\ell k}^{(a)}
   \sum_{b=0}^q \p{\ell-k}^{q-b} { q \choose b }
@@
 
-In this section we work out the effects of decoherence via
-collective-spin jump operators on ordered products of collective spin
-operators.  For shorthand, we define
+In this section we work out the effects of decoherence via collective
+jump operators on ordered products of collective spin operators.  For
+shorthand, we define
 \begin{align}
@@
 \end{align}
-where $\Gamma$ is a collective spin operator.
+where $\Gamma$ is a collective spin operator with weight one i.e.~a
+linear combination of $S_\z$, $S_\mu$, and $S_\bmu$.
 
@@
 \end{align}
-using which we can compute, for $\bmu\equiv-\mu\in\set{+1,-1}$,
-\begin{align}
-  \bk{\pm\Z|S_\mu^\ell S_\z^m S_\bmu^n|\pm\Z}
-  = \delta_{\ell,0} \delta_{m,0} \p{\pm S}^m,
+using which we can compute, for $\bmu\equiv-\mu\in\set{+1,-1}$ and
+$\bnu\equiv-\nu\in\set{+1,-1}$,
+\begin{align}
+  \bk{\nu\Z|S_\mu^\ell S_\z^m S_\bmu^n|\nu\Z}
+  = \delta_{\ell n} \nu^m \times
+  \begin{cases}
+    \delta_{n,0} S^m & \mu \ne \nu, \\
+    \p{S-n}^m \f{N! n!}{\p{N-n}!} & \mu = \nu,
+  \end{cases}
 \end{align}
@@
   ^{S-\delta_{\mu,1}\max\set{\ell,n}} k^m
-  \f{\p{S-\mu k}!}{\p{S+\mu k}!\p{S-\mu k-\ell}!\p{S-\mu k-n}!}.
-\end{align}
-% With the help of Mathematica, we can write this result in the
-% alternate form
-% \begin{align}
-%   \bk{\X|S_+^\ell S_\z^m S_-^n|\X}
-%   = \f{\p{N!}^2}{2^N\p{N-\ell}!\p{N-n}!}
-%   \sp{\p{-S}^m \alpha + \f{\p{N-\ell}\p{N-n}}{N} \beta},
-% \end{align}
-% where $\alpha$ and $\beta$ are defined in terms of the generalized
-% hypergeometric function $\F$ by
-% \begin{align}
-%   \alpha
-%   \equiv \F\p{\let\scriptstyle\textstyle
-%     \substack{\ell-N,~n-N \\ -N}~; -1},
-% \end{align}
-% and
-% \begin{align}
-%   \beta
-%   \equiv \sum_{p=0}^{m-1} \p{-S}^p { m \choose p }
-%   \tilde\F_{m-p-1}\p{\let\scriptstyle\textstyle
-%     \substack{1+\ell-N,~1+n-N \\ 1-N}~; -1},
-% \end{align}
-% for
-% \begin{align}
-%   \tilde\F_q\p{\let\scriptstyle\textstyle
-%     \substack{a,~b\\c}~; z}
-%   \equiv \F\p{\let\scriptstyle\textstyle
-%     \substack{2,2,\cdots,2,a,b \\ 1,1,\cdots,1,c}~; z}
-% \end{align}
-% with $q$ repetitions of 2 and 1.
+  \f{\p{S-\mu k}!}{\p{S+\mu k}!\p{S-\mu k-\ell}!\p{S-\mu k-n}!} \\
+  &= \p{-\mu}^m \f{N!}{2^N} \sum_{k=-S+\max\set{\ell,n}}^S k^m
+  \f{\p{S+k}!}{\p{S-k}!\p{S+k-\ell}!\p{S+k-n}!}.
+\end{align}
+Furthermore, using the fact that
+\begin{align}
+  e^{i\phi S_\z} S_\mu e^{-i\phi S_\z} = e^{i\mu\phi} S_\mu,
+\end{align}
+we can say
+\begin{align}
+  \bk{\phi|S_\mu^\ell S_\z^m S_\bmu^n|\phi}
+  = e^{i\mu\p{\ell-n}\phi} \bk{\X|S_\mu^\ell S_\z^m S_\bmu^n|\X},
+\end{align}
+where
+\begin{align}
+  \ket{\phi} \equiv e^{-i\phi S_\z}\ket{\X},
+  &&
+  \p{\cos\phi S_\x + \sin\phi S_\y} \ket\phi = S \ket\phi.
+\end{align}
 

### 2018-09-11 soc_squeezing/scripts/correlator_methods.py

@@ -3,7 +3,6 @@
 # FILE CONTENTS: methods for computing collective spin correlators
 
 import numpy as np
-import scipy.sparse as sparse
 import itertools
 
 from math import lgamma
@@ -17,20 +16,17 @@ from sympy.functions.combinatorial.numbers import stirling as sympy_stirling
 ##########################################################################################
 
 # natural logarithm of factorial
-def ln_factorial(n): return lgamma(n+1)
+def ln_factorial(nn, ln_factorials = {}):
+    try: return ln_factorials[nn]
+    except:
+        ln_factorials[nn] = lgamma(nn+1)
+        return ln_factorials[nn]
 
-def ln_factors_pX(total_spin, kk, ll, nn, ln_factorials = {}):
-    ln_term = 0
-    for fac_val, sign in [ (total_spin + kk, 1),
-                           (total_spin - kk, -1),
-                           (total_spin + kk - ll, -1),
-                           (total_spin + kk - nn, -1) ]:
-        try:
-            ln_term += sign * ln_factorials[fac_val]
-        except:
-            ln_factorials[fac_val] = ln_factorial(fac_val)
-            ln_term += sign * ln_factorials[fac_val]
-    return ln_term
+def ln_factors_pX(total_spin, kk, ll, nn):
+    return ( ln_factorial(total_spin+kk)
+             - ln_factorial(total_spin - kk)
+             - ln_factorial(total_spin + kk - ll)
+             - ln_factorial(total_spin + kk - nn) )
 ln_factors_pX = np.vectorize(ln_factors_pX)
 
 # correlator < +X | S_\mu^ll S_\z^mm S_\nu^nn | +X >
@@ -115,15 +111,13 @@ def ext_binom_op(ll, mm, nn, x, terms, prefactor = 1):
 ##########################################################################################
 
 # unsigned stirling number of the first kind
-def stirling(n,k): return float(sympy_stirling(n, k, kind = 1, signed = True))
+def stirling(nn,kk): return float(sympy_stirling(nn, kk, kind = 1, signed = False))
 
-# collective spin operator commutator coefficients (see notes)
-def epsilon(mm,nn,pp,ll):
[…]

### 2018-09-11 soc_squeezing/cumulant_expansions.tex
aesthetic changes

@@
 \newcommand{\bmu}{{\bar\mu}}
-\newcommand{\bnu}{{\bar\nu}}
 \newcommand{\floor}[1]{\left\lfloor{#1}\right\rfloor}
@@
 \begin{align}
-  \Phi\p{\zeta,t}
-  &\equiv e^{-\lambda t} \sp{\cos\p{t\sqrt{s_\zeta^2-r}}
-    + \lambda t~\t{sinc}\p{t\sqrt{s_\zeta^2-r}}},
+  \Phi\p{X,t}
+  &\equiv e^{-\lambda t} \sp{\cos\p{t\sqrt{s_X^2-r}}
+    + \lambda t~\t{sinc}\p{t\sqrt{s_X^2-r}}},
   \\
-  \Psi\p{\zeta,t}
-  &\equiv e^{-\lambda t} \p{is_\zeta-\gamma}t~
-  \t{sinc}\p{t\sqrt{s_\zeta^2-r}},
+  \Psi\p{X,t}
+  &\equiv e^{-\lambda t} \p{is_X-\gamma}t~
+  \t{sinc}\p{t\sqrt{s_X^2-r}},
 \end{align}
@@
   &&
-  s_\zeta \equiv \zeta + i\gamma,
+  s_X \equiv X + i\gamma,
 \end{align}
@@
   S_\bmu^{n-p} S_\mu^{m-p} \sum_{\ell=0}^p
-  \mu^\ell \epsilon_{mnp}^{(\ell)} S_\z^\ell,
+  \mu^\ell \epsilon_{mn}^{p\ell} S_\z^\ell,
 \end{align}
@@
 \begin{align}
-  \epsilon_{mnp}^{(\ell)}
+  \epsilon_{mn}^{p\ell}
   \equiv 2^\ell \sum_{q=\ell}^p \p{-1}^{p-q}
   { p \brack q } { q \choose \ell } \p{m-n}^{q-\ell},
-  \label{eq:eps_mnp}
 \end{align}
@@
   p! { m \choose p } { n \choose p } S_\bmu^{n-p} S_\mu^{m-p}
-  \sum_{\ell=0}^p \mu^\ell \epsilon_{mnp}^{(\ell)} S_\z^\ell,
+  \sum_{\ell=0}^p \mu^\ell \epsilon_{mn}^{p\ell} S_\z^\ell,
   \label{eq:comm_mu}
@@
   p! { m \choose p } { n \choose p } S_\bmu^{n-p} S_\mu^{m-p}
-  \sum_{\ell=0}^p \mu^\ell \epsilon_{mnp}^{(\ell)} S_\z^\ell,
+  \sum_{\ell=0}^p \mu^\ell \epsilon_{mn}^{p\ell} S_\z^\ell,
   \label{eq:push_mu_bmu}
@@
   Z_{mn;\mu}^{(p)}
-  \equiv \sum_{\ell=0}^p \mu^\ell \epsilon_{mnp}^{(\ell)}
+  \equiv \sum_{\ell=0}^p \mu^\ell \epsilon_{mn}^{p\ell}
   \p{-\mu\sp{m-p} + S_\z}^\ell
-  = \sum_{q=0}^p \mu^q \zeta_{mnp}^{(q)} S_\z^q,
-  \label{eq:Z_mn}
+  = \sum_{q=0}^p \mu^q \zeta_{mn}^{pq} S_\z^q,
 \end{align}
@@
 \begin{align}
-  \zeta_{mnp}^{(q)}
+  \zeta_{mn}^{pq}
   \equiv \sum_{\ell=q}^p \p{-1}^{\ell-q}
-  \epsilon_{mnp}^{(\ell)} { \ell \choose q } \p{m-p}^{\ell-q}
+  \epsilon_{mn}^{p\ell} { \ell \choose q } \p{m-p}^{\ell-q}
   = \p{-1}^{p-q} 2^q \sum_{s=q}^p
   { p \brack s } { s \choose q } \p{m+n-2p}^{s-q}.
-  \label{eq:zeta_mnp}
+  \label{eq:zeta_mnpq}
 \end{align}
@@
 \end{align}
-and, in terms of $\zeta_{r\ell k}^{(a)}$ as defined in
-\eqref{eq:zeta_mnp},
+and, in terms of $\zeta_{r\ell}^{ka}$ as defined in
+\eqref{eq:zeta_mnpq},
 \begin{align}
@@
   Z_{r\ell;\bmu}^{(k)} \p{\mu\sp{r-k}+S_\z}^m \\
-  &= \sum_{a=0}^k \p{-1}^a \zeta_{r\ell k}^{(a)}
+  &= \sum_{a=0}^k \p{-1}^a \zeta_{r\ell}^{ka}
   \sum_{b=0}^q \p{\ell-k}^{q-b} { q \choose b }
@@
 \end{align}
+Note that the sum $\ell+m+n$ does not ever grow under this form of
+decoherence.
 
@@
 \end{align}
+Note that the sum $\ell+m+n$ grows by one if $\Gamma_\mu\ne0$ or
+$\Gamma_\bmu\ne0$, and does not grow otherwise.
 
@@
 \end{align}
-using which we can compute, for $\bmu\equiv-\mu\in\set{+1,-1}$ and
-$\bnu\equiv-\nu\in\set{+1,-1}$,
+using which we can compute, for $\mu,\nu\in\set{+1,-1}$ and
+$\bmu\equiv-\mu$,
 \begin{align}
@@
   \begin{cases}
-    \delta_{n,0} S^m & \mu \ne \nu, \\
-    \p{S-n}^m \f{N! n!}{\p{N-n}!} & \mu = \nu,
+     \p{S-n}^m \f{N! n!}{\p{N-n}!} & \mu = \nu, \\
+     \delta_{n,0} S^m & \mu \ne \nu,
   \end{cases}

### 2018-09-12 soc_squeezing/cumulant_expansions.tex
add text explaining transformation of spin operators for TAT protocol

@@
 
-\section{One-axis twisting revisited}
+\section{Decoherence in the two-axis twisting protocols}
+
+In the case of squeezing through spin-orbit coupling with the Sr-87
+optical lattice clock, the primary form of decoherence is through
+decay of individual spins, i.e.~with jump operators $\sigma_-^{(j)}$.
+The two axis twisting Hamiltonians $H_{\t{TAT}}^{\z,\y}$ and
+$H_{\t{TAT}}^{\z,\y}$, however, are actually realized in a rotating
+frame, which implies that we must transform these jump operators
+appropriately.  Specifically, the TAT protocol begins with a
+time-dependent transverse-field Hamiltonian
+\begin{align}
+  H = \chi S_\z^2 + \Omega\p{t} S_\x,
+  &&
+  \Omega\p{t} = \beta\omega\cos\p{\omega t} S_\x,
+\end{align}
+where $\beta$ is the modulation index of the driving field, and the
+drive frequency $\omega\gg N\chi$.  Moving into the rotating frame of
+$\Omega\p{t}S_\x$ subtracts this term from the Hamiltonian, and
+transforms operators $\O$ as
+\begin{align}
+  \O \to U\p{t}^\dag \O U\p{t},
+  &&
+  U\p{t} \equiv \exp\p{-i\int_0^t\d\tau~\Omega\p{\tau}S_\x}
+  = \exp\sp{-i\beta\sin\p{\omega t}S_\x},
+\end{align}
+In particular, the operators $\tilde S_\pm\equiv-S_\z\pm iS_\y$
+transform as
+\begin{align}
+  \tilde S_\pm
+  \to U^\dag \tilde S_\pm U
+  = e^{\pm i\beta\sin\p{\omega t}} \tilde S_\pm.
+\end{align}
+For large drive frequencies $\omega$, we can generally make a secular
+approximation to say
+\begin{align}
+  e^{\pm i\beta\sin\p{\omega t}}
+  = \sum_{n=-\infty}^\infty J_n\p{\pm\beta} e^{in\omega t}
+  \approx J_0\p{\pm\beta} = J_0\p{\beta},
+\end{align}
+where $J_0$ is the zero-order Bessel function of the first kind.
+Expanding $S_\z^2=\f14\p{\tilde S_++\tilde S_-}^2$, one can thus work
+out that the effective Hamiltonian in the rotating frame of the drive
+is
+\begin{align}
+  H_{\t{eff}}
+  \approx \f{\chi}{2} \p{\sp{J_0\p{2\beta}+1} S_\z^2
+  - \sp{J_0\p{2\beta}-1} S_\y^2},
+\end{align}
+where choosing $\beta$ such that $J_0\p{2\beta}=\pm1/3$ will give us
+$H_{\t{TAT}}^{\z,\y}$ or $H_{\t{TAT}}^{\y,\x}$ up to irrelevant terms.
+In a similar spirit, one can work out that
+\begin{align}
+  \tilde\sigma_\pm^{(j)}
+  \equiv \f12\p{-\sigma_\z^{(j)}\pm i\sigma_\y^{(j)}}
+  \to U^\dag \tilde\sigma_\pm^{(j)} U
+  = e^{\pm i\beta\sin\p{\omega t}} \tilde\sigma_\pm^{(j)}
+  \approx J_0\p{\beta} \tilde\sigma_\pm^{(j)},
+\end{align}
+which implies
+\begin{align}
+  \sigma_\z^{(j)} \to J_0\p{\beta} \sigma_\z^{(j)},
+  &&
+  \sigma_\pm^{(j)}
+  \to \f12\sp{1 \pm J_0\p{\beta}} \sigma_+^{(j)}
+  + \f12\sp{1 \mp J_0\p{\beta}} \sigma_-^{(j)}.
+\end{align}
+The TAT protocol about $\p{\z,\y}$ requires
+\begin{align}
+  J_0\p{2\beta} = 1/3
+  \implies \beta \approx 0.906
+  \implies J_0\p{\beta} \approx 0.805,
+\end{align}
+while the TAT protocol about $\p{\y,\x}$ requires
+\begin{align}
+  J_0\p{2\beta} = -1/3
+  \implies
+  \begin{cases}
+    \beta \approx 1.63
+    \implies J_0\p{\beta} \approx 0.440, \\
+    \beta \approx 2.22
+    \implies J_0\p{\beta} \approx 0.099.
+  \end{cases}
+\end{align}
 

### 2018-09-13 soc_squeezing/scripts/correlator_methods.py
fixed small errors in notes and code comments

@@ -224,7 +224,7 @@ def multiply_vecs(vec_left, vec_right, mu, prefactor = 1):
 # miscellaneous methods for changing frames and operator vectors
 ##########################################################################################
 
-# decoherence transformation matrix from a periodic drive; here A = J_0(2\beta), where:
+# decoherence transformation matrix from a periodic drive; here A = J_0(\beta), where:
 #   J_0 is the zero-order bessel function of the first kind
 #   \beta is the modulation index
 def dec_mat_drive(A):

### 2018-09-13 soc_squeezing/cumulant_expansions.tex
fixed small errors in notes and code comments
minor clarification in notes
added back the expressions for expectation values in terms of hypergeometric functions

@@
 \end{align}
-where Hermiticity requires that $h_{\ell mn}=h_{nm\ell}$.  The
+where Hermiticity requires that $h_{\ell mn}=h_{nm\ell}^*$.  The
 Hamiltonian for one-axis twisting with a transverse field, for
@@
 \end{align}
-where $J_0$ is the zero-order Bessel function of the first kind.
+where $J_n$ is the $n$-th order Bessel function of the first kind.
 Expanding $S_\z^2=\f14\p{\tilde S_++\tilde S_-}^2$, one can thus work
@@
 \end{align}
+Alternatively, with the help of Mathematica we can find
+\begin{align}
+  \bk{\X|S_\mu^\ell S_\z^m S_\bmu^n|\X}
+  = \f{\mu^m\p{N!}^2}{2^N\p{N-\ell}!\p{N-n}!}
+  \sp{\p{-S}^m \alpha + \f{\p{N-\ell}\p{N-n}}{N} \beta},
+\end{align}
+where $\alpha$ and $\beta$ are defined in terms of the generalized
+hypergeometric function $\F$ by
+\begin{align}
+  \alpha
+  \equiv \F\p{\let\scriptstyle\textstyle
+    \substack{\ell-N,~n-N \\ -N}~; -1},
+\end{align}
+and
+\begin{align}
+  \beta
+  \equiv \sum_{p=0}^{m-1} \p{-S}^p { m \choose p }
+  \tilde\F_{m-p-1}\p{\let\scriptstyle\textstyle
+    \substack{1+\ell-N,~1+n-N \\ 1-N}~; -1},
+\end{align}
+for
+\begin{align}
+  \tilde\F_q\p{\let\scriptstyle\textstyle
+    \substack{a,~b\\c}~; z}
+  \equiv \F\p{\let\scriptstyle\textstyle
+    \substack{2,2,\cdots,2,a,b \\ 1,1,\cdots,1,c}~; z}
+\end{align}
+with $q$ repetitions of 2 and 1.
 

### 2018-09-14 soc_squeezing/cumulant_expansions.tex
make S_z always appear as mu S_z

@@
 \newcommand{\bmu}{{\bar\mu}}
+\newcommand{\bnu}{{\bar\nu}}
 \newcommand{\floor}[1]{\left\lfloor{#1}\right\rfloor}
@@
   \f12 \sp{\sigma_\z^{(j)},\sigma_\mu^{(k)}}_-
-  &= \delta_{jk} \mu \sigma_\mu^{(j)},
+  &= \delta_{jk} \mu\sigma_\mu^{(j)},
   &
   \sp{S_\z,\sigma_\mu^{(j)}}_-
-  &= \f12 \sp{\sigma_\z^{(j)},S_\mu}_- = \mu \sigma_\mu^{(j)},
+  &= \f12 \sp{\sigma_\z^{(j)},S_\mu}_- = \mu\sigma_\mu^{(j)},
   &
@@
 \begin{align}
-  S_\z^m \sigma_\mu^{(j)}
-  = S_\z^{m-1} \sigma_\mu^{(j)} \p{\mu + S_\z}
-  = S_\z^{m-2} \sigma_\mu^{(j)} \p{\mu + S_\z}^2
+  \p{\mu S_\z}^m \sigma_\mu^{(j)}
+  = \p{\mu S_\z}^{m-1} \sigma_\mu^{(j)} \p{1 + \mu S_\z}
+  = \p{\mu S_\z}^{m-2} \sigma_\mu^{(j)} \p{1 + \mu S_\z}^2
   = \cdots
-  = \sigma_\mu^{(j)} \p{\mu + S_\z}^m,
+  = \sigma_\mu^{(j)} \p{1 + \mu S_\z}^m,
   \label{eq:push_z_mu_Ss}
@@
 \begin{align}
-  \sigma_\z^{(j)} S_\mu^m
-  = \p{2\mu\sigma_\mu^{(j)} + S_\mu \sigma_\z^{(j)}} S_\mu^{m-1}
-  = 2\mu\sigma_\mu^{(j)} S_\mu^{m-1} + S_\mu \sigma_\z^{(j)} S_\mu^{m-1}
+  \mu\sigma_\z^{(j)} S_\mu^m
+  = S_\mu \mu\sigma_\z^{(j)} S_\mu^{m-1} + 2\sigma_\mu^{(j)} S_\mu^{m-1}
   = \cdots
-  = 2\mu m\sigma_\mu^{(j)} S_\mu^{m-1} + S_\mu^m \sigma_\z^{(j)}.
+  = 2m\sigma_\mu^{(j)} S_\mu^{m-1} + S_\mu^m \mu \sigma_\z^{(j)},
   \label{eq:push_z_mu_sS}
 \end{align}
-Summing over the single-spin index $j$ in both of these cases gives us
-the purely collective-spin versions of these identities:
-\begin{align}
-  S_\z^m S_\mu = S_\mu \p{\mu + S_\z}^m,
-  &&
-  S_\z S_\mu^m = S_\mu^m \p{\mu m + S_\z},
+where we will generally find it nicer to express our results in terms
+of $\mu\sigma_\z^{(j)}$ and $\mu S_\z$ rather than $\sigma_\z^{(j)}$
+and $S_\z$.  Summing over the single-spin index $j$ in both of these
+cases gives us the purely collective-spin versions of these
+identities:
+\begin{align}
+  \p{\mu S_\z}^m S_\mu = S_\mu \p{1 + \mu S_\z}^m,
+  &&
+  \mu S_\z S_\mu^m = S_\mu^m \p{m + \mu S_\z},
   \label{eq:push_z_mu_single}
@@
 \begin{align}
-  S_\z^\ell S_\mu^m
-  = S_\z^{\ell-1} S_\mu^m \p{\mu m + S_\z}
-  = S_\z^{\ell-2} S_\mu^m \p{\mu m + S_\z}^2
+  \p{\mu S_\z}^\ell S_\mu^m
+  = \p{\mu S_\z}^{\ell-1} S_\mu^m \p{m + \mu S_\z}
+  = \p{\mu S_\z}^{\ell-2} S_\mu^m \p{m + \mu S_\z}^2
   = \cdots
-  = S_\mu^m \p{\mu m + S_\z}^\ell.
+  = S_\mu^m \p{m + \mu S_\z}^\ell.
   \label{eq:push_z_mu}
 \end{align}
-Similarly finding commutation relations between powers of transverse
-spin operators, i.e.~$S_\mu$, turns out to be considerably more
-difficult than the cases we have worked out thus far.  We therefore
-save this work for Appendix \ref{sec:comm_transverse}.
+Multiplying \eqref{eq:push_z_mu} through by $\p{\mu\nu}^\ell$ (for
+$\nu\in\set{+1,-1}$) and taking its Hermitian conjugate, we can say
+that more generally
+\begin{align}
+  \p{\nu S_z}^\ell S_\mu^m
+  = S_\mu^m \p{\mu\nu m+\nu S_\z}^\ell,
+  &&
+  S_\mu^m \p{\nu S_z}^\ell
+  = \p{-\mu\nu m+\nu S_\z}^\ell S_\mu^m.
+\end{align}
+Finding commutation relations between powers of transverse spin
+operators, i.e.~$S_\mu$ and $S_\bmu$, turns out to be considerably
+more difficult than the cases we have worked out thus far.  We
+therefore save this work for Appendix \ref{sec:comm_transverse}.
 
@@
   &= S_\mu^{m-1}\sigma_\bmu^{(j)} S_\mu
-  + \mu S_\mu^{m-1} \sigma_\z^{(j)} \\
+  + S_\mu^{m-1} \mu\sigma_\z^{(j)} \\
   &= S_\mu^{m-2} \sigma_\bmu^{(j)} S_\mu^2
-  + \mu S_\mu^{m-2} \sigma_\z^{(j)} S_\mu
-  + \mu S_\mu^{m-1} \sigma_\z^{(j)} \\
+  + S_\mu^{m-2} \mu\sigma_\z^{(j)} S_\mu
+  + S_\mu^{m-1} \mu\sigma_\z^{(j)} \\
   &= \sigma_\bmu^{(j)} S_\mu^m
-  + \mu \sum_{k=0}^{m-1} S_\mu^k \sigma_\z^{(j)} S_\mu^{m-k-1}
+  + \sum_{k=0}^{m-1} S_\mu^k \mu\sigma_\z^{(j)} S_\mu^{m-k-1}
   \label{eq:push_mu_Ss_start}.
@@
 ordering on products of spin operators, which will allow us to only
-keep track of operators that are linearly independent.  We choose to
-impose an ordering with all $\sigma_\bmu^{(j)}$ operators on the left,
-and all $\sigma_\z^{(j)}$ operators on the right, as such an ordering
-will prove convenient for the calculations in this section.  This
-choice of ordering compels us to expand
-\begin{align}
-  \mu \sum_{k=0}^{m-1} S_\mu^k \sigma_\z^{(j)} S_\mu^{m-k-1}
-  &= \mu \sum_{k=0}^{mq-1} S_\mu^k
-  \sp{2\mu\p{m-k-1} \sigma_\mu^{(j)} S_\mu^{m-k-2}
-    + S_\mu^{m-k-1} \sigma_\z^{(j)}} \\
+keep track of operators that are linearly independent.  We choose (for
+now) to impose an ordering with all $\sigma_\bmu^{(j)}$ operators on
+the left, and all $\sigma_\z^{(j)}$ operators on the right, as such an
+ordering will prove convenient for the calculations in this section.
+This choice of ordering compels us to expand
+\begin{align}
+  \sum_{k=0}^{m-1} S_\mu^k \mu\sigma_\z^{(j)} S_\mu^{m-k-1}
+  &= \sum_{k=0}^{m-1} S_\mu^k
+  \sp{2\p{m-k-1} \sigma_\mu^{(j)} S_\mu^{m-k-2}
+    + S_\mu^{m-k-1} \mu\sigma_\z^{(j)}} \\
   &= m \p{m-1} \sigma_\mu^{(j)} S_\mu^{m-2}
-  + \mu m S_\mu^{m-1} \sigma_\z^{(j)},
+  + m S_\mu^{m-1} \mu\sigma_\z^{(j)},
 \end{align}
@@
   = \sigma_\bmu^{(j)} S_\mu^m + m \p{m-1} \sigma_\mu^{(j)} S_\mu^{m-2}
-  + \mu m S_\mu^{m-1} \sigma_\z^{(j)},
+  + m S_\mu^{m-1} \mu\sigma_\z^{(j)},
   \label{eq:push_mu_Ss}
@@
   { q \choose \ell } \p{m-n}^{q-\ell} \p{2\mu S_\z}^\ell \\
-  &= \sum_{\ell=0}^p \p{2\mu}^\ell \sum_{q=\ell}^p \p{-1}^{p-q}
-  { p \brack q } { q \choose \ell } \p{m-n}^{q-\ell} S_\z^\ell.
+  &= \sum_{\ell=0}^p 2^\ell \sum_{q=\ell}^p \p{-1}^{p-q}
+  { p \brack q } { q \choose \ell } \p{m-n}^{q-\ell} \p{\mu S_\z}^\ell.
 \end{align}
@@
   = p! { m \choose p } { n \choose p }
-  S_\bmu^{n-p} S_\mu^{m-p} \sum_{\ell=0}^p
-  \mu^\ell \epsilon_{mn}^{p\ell} S_\z^\ell,
+  S_\bmu^{n-p} S_\mu^{m-p}
+  \sum_{\ell=0}^p \epsilon_{mn}^{p\ell} \p{\mu S_\z}^\ell,
 \end{align}
@@
   p! { m \choose p } { n \choose p } S_\bmu^{n-p} S_\mu^{m-p}
-  \sum_{\ell=0}^p \mu^\ell \epsilon_{mn}^{p\ell} S_\z^\ell,
+  \sum_{\ell=0}^p \epsilon_{mn}^{p\ell} \p{\mu S_\z}^\ell,
   \label{eq:comm_mu}
@@
   p! { m \choose p } { n \choose p } S_\bmu^{n-p} S_\mu^{m-p}
-  \sum_{\ell=0}^p \mu^\ell \epsilon_{mn}^{p\ell} S_\z^\ell,
+  \sum_{\ell=0}^p \epsilon_{mn}^{p\ell} \p{\mu S_\z}^\ell,
   \label{eq:push_mu_bmu}
@@
   = \sum_{p=0}^{\min\set{m,n}} p! { m \choose p } { n \choose p }
-  S_\bmu^{n-p} Z_{mn;\mu}^{(p)} S_\mu^{m-p},
+  S_\bmu^{n-p} Z_{mn;\bmu}^{(p)} S_\mu^{m-p},
 \end{align}
@@
 \begin{align}
-  Z_{mn;\mu}^{(p)}
-  \equiv \sum_{\ell=0}^p \mu^\ell \epsilon_{mn}^{p\ell}
-  \p{-\mu\sp{m-p} + S_\z}^\ell
-  = \sum_{q=0}^p \mu^q \zeta_{mn}^{pq} S_\z^q,
+  Z_{mn;\bmu}^{(p)}
+  \equiv \sum_{\ell=0}^p \epsilon_{mn}^{p\ell}
+  \p{-\sp{m-p} + \mu S_\z}^\ell
+  = \sum_{\ell=0}^p \epsilon_{mn}^{p\ell}
+  \p{-1}^\ell \p{\sp{m-p} + \bmu S_\z}^\ell
+  = \sum_{q=0}^p \zeta_{mn}^{pq} \p{\bmu S_\z}^q,
+  \label{eq:Z_mnp}
 \end{align}
@@
   \zeta_{mn}^{pq}
-  \equiv \sum_{\ell=q}^p \p{-1}^{\ell-q}
-  \epsilon_{mn}^{p\ell} { \ell \choose q } \p{m-p}^{\ell-q}
-  = \p{-1}^{p-q} 2^q \sum_{s=q}^p
+  \equiv \sum_{\ell=q}^p \epsilon_{mn}^{p\ell} \p{-1}^\ell
+  { \ell \choose q } \p{m-p}^{\ell-q}
+  = \p{-1}^p 2^q \sum_{s=q}^p
   { p \brack s } { s \choose q } \p{m+n-2p}^{s-q}.
@@
 
-If all collective spin operators are written in the form
-$S_\mu^\ell S_\z^m S_\bmu^n$ for $\bmu\equiv-\mu\in\set{+1,-1}$, then
-the most general (anti-)commutator we need to consider is
-\begin{align}
-  \sp{S_\mu^p S_\z^q S_\bmu^r, S_\mu^\ell S_\z^m S_\bmu^n}_\pm
+We will generally be concerned with ordered products of collective
+spin operators of the form $S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$ for
+$\bmu\equiv-\mu\in\set{+1,-1}$.  The most general (anti-)commutator we
+need to consider is then
+\begin{align}
+  \sp{S_\mu^p \p{\mu S_\z}^q S_\bmu^r,
+    S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_\pm
   = \S^{pqr}_{\ell mn;\mu} \pm \S^{\ell mn}_{pqr;\mu},
@@
   \S^{pqr}_{\ell mn;\mu}
-  \equiv S_\mu^p S_\z^q S_\bmu^r S_\mu^\ell S_\z^m S_\bmu^n
+  = S_\mu^p \p{\mu S_\z}^q S_\bmu^r
+  S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n
   = \sum_{k=0}^{\min\set{r,\ell}} k! { r \choose k } { \ell \choose k }
@@
 \end{align}
-and, in terms of $\zeta_{r\ell}^{ka}$ as defined in
-\eqref{eq:zeta_mnpq},
+for
 \begin{align}
   \tilde Z_{qr\ell m;\mu}^{(k)}
-  &\equiv \p{\mu\sp{\ell-k}+S_\z}^q
-  Z_{r\ell;\bmu}^{(k)} \p{\mu\sp{r-k}+S_\z}^m \\
-  &= \sum_{a=0}^k \p{-1}^a \zeta_{r\ell}^{ka}
+  &\equiv \p{\sp{\ell-k}+\mu S_\z}^q
+  Z_{r\ell;\mu}^{(k)} \p{\sp{r-k}+\mu S_\z}^m \\
+  &= \sum_{a=0}^k \zeta_{r\ell}^{ka}
   \sum_{b=0}^q \p{\ell-k}^{q-b} { q \choose b }
   \sum_{c=0}^m \p{r-k}^{m-c} { m \choose c }
-  \mu^{q+m+a-b-c} S_\z^{a+b+c}.
-\end{align}
+  \p{\mu S_\z}^{a+b+c},
+\end{align}
+defined in terms of $Z_{r\ell;\mu}^{(k)}$ and $\zeta_{r\ell}^{ka}$ as
+given in \eqref{eq:Z_mnp} and \eqref{eq:zeta_mnpq}.
 
@@
 
-We first compute sums of single-spin operators sandwiching $S_\z^m$,
-which will in turn be used to compute similar sums sandwiching
-$S_\mu^\ell S_\z^m S_\bmu^n$.  The simplest case is
-\begin{align}
-  \sum_j \sigma_\z^{(j)} S_\z^m \sigma_\z^{(j)}
-  = \sum_j \sigma_\z^{(j)} \sigma_\z^{(j)} S_\z^m
-  = \sum_j \1_j S_\z^m
-  = 2 S S_\z^m.
-\end{align}
-Similarly,
-\begin{align}
-  \sum_j \sigma_\z^{(j)} S_\z^m \sigma_\mu^{(j)}
-  = S_\z^m \sum_j \sigma_\z^{(j)} \sigma_\mu^{(j)}
-  = S_\z^m \sum_j \mu \sigma_\mu^{(j)}
-  = \mu S_\z^m S_\mu
-  = \mu S_\mu \p{\mu+S_\z}^m,
-\end{align}
-where in the last equality we used the identity derived in
-\eqref{eq:push_z_mu_single} to push $S_\mu$ through $S_\z^m$.  Using
-identity \eqref{eq:push_z_mu_Ss} to push $\sigma_\mu^{(j)}$ through
-$S_\z^m$, we can also find
-\begin{align}
-  \sum_j \sigma_\mu^{(j)} S_\z^m \sigma_\mu^{(j)}
-  = \sum_j \sigma_\mu^{(j)} \sigma_\mu^{(j)} \p{\mu+S_\z}^m
-  = 0,
-\end{align}
-and
-\begin{align}
-  \sum_j \sigma_\bmu^{(j)} S_\z^m \sigma_\mu^{(j)}
-  = \sum_j \sigma_\bmu^{(j)} \sigma_\mu^{(j)} \p{\mu+S_\z}^m
-  = \p{S-\mu S_\z} \p{\mu+S_\z}^m.
+We first compute sums of single-spin operators sandwiching
+$\p{\mu S_\z}^m$, which will in turn be used to compute similar sums
+sandwiching $S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$.  Up to Hermitian
+conjugation, the unique cases are, for $\mu,\nu\in\set{+1,-1}$ and
+$\bmu\equiv-\mu$,
+\begin{align}
+  \sum_j \sigma_\z^{(j)} \p{\mu S_\z}^m \sigma_\z^{(j)}
+  &= \sum_j \sigma_\z^{(j)} \sigma_\z^{(j)} \p{\mu S_\z}^m
+  = \sum_j \1_j \p{\mu S_\z}^m
+  = 2 S \p{\mu S_\z}^m, \\
+  \sum_j \sigma_\z^{(j)} \p{\mu S_\z}^m \sigma_\nu^{(j)}
+  &= \p{\mu S_\z}^m \sum_j \sigma_\z^{(j)} \sigma_\nu^{(j)}
+  = \p{\mu S_\z}^m \nu S_\nu
+  = \nu S_\nu \p{\mu\nu+\mu S_\z}^m, \\
+  \sum_j \sigma_\nu^{(j)} \p{\mu S_\z}^m \sigma_\nu^{(j)}
+  &= \sum_j \sigma_\nu^{(j)} \sigma_\nu^{(j)} \p{\mu\nu+\mu S_\z}^m
+  = 0, \\
+  \sum_j \sigma_\bnu^{(j)} \p{\mu S_\z}^m \sigma_\nu^{(j)}
+  &= \sum_j \sigma_\bnu^{(j)} \sigma_\nu^{(j)} \p{\mu\nu+\mu S_\z}^m
+  = \p{S-\nu S_\z} \p{\mu\nu+\mu S_\z}^m.
 \end{align}
 We are now equipped to derive similar identities for general ordered
-products of collective spin operators.  Making (heavy) use of
-identities \eqref{eq:push_z_mu_sS} and \eqref{eq:push_mu_Ss} to push
-single-spin operators through transverse collective-spin operators, we
-work through all combinations that are unique up to Hermitian
-conjugation, finding
-\begin{align}
-  \sum_j \sigma_\z^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\z^{(j)}
-  &= 2\p{S-\ell-n} S_\mu^\ell S_\z^m S_\bmu^n
-  + 4\ell n S_\mu^{\ell-1} \p{S+\mu S_\z}
-  \p{-\mu+S_\z}^m S_\bmu^{n-1},
+products of collective spin operators.  Making heavy use of identities
+\eqref{eq:push_z_mu_sS} and \eqref{eq:push_mu_Ss} to push single-spin
+operators through transverse collective-spin operators, we again work
+through all combinations that are unique up to Hermitian conjugation,
+finding
+\begin{align}
+  \sum_j \sigma_\z^{(j)} S_\mu^\ell
+  \p{\mu S_\z}^n S_\bmu^n \sigma_\z^{(j)}
+  &= 2\p{S-\ell-n} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n \notag \\
+  &\qquad + 4\ell n S_\mu^{\ell-1} \p{S+\mu S_\z}
+  \p{-1+\mu S_\z}^m S_\bmu^{n-1},
   \label{eq:san_z_z} \allowdisplaybreaks \\
-  \sum_j \sigma_\z^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\mu^{(j)}
-  &= \mu S_\mu^{\ell+1} \p{\mu+S_\z}^m S_\bmu^n
-  + \mu n \p{-2S+2\ell+n-1} S_\mu^\ell S_\z^m S_\bmu^{n-1} \notag \\
+  \sum_j \sigma_\z^{(j)} S_\mu^\ell
+  \p{\mu S_\z}^n S_\bmu^n \sigma_\mu^{(j)}
+  &= \mu S_\mu^{\ell+1} \p{1+\mu S_\z}^m S_\bmu^n
+  + \mu n \p{-2S+2\ell+n-1} S_\mu^\ell
+  \p{\mu S_\z}^m S_\bmu^{n-1} \notag \\
   &\qquad - 2\mu\ell n\p{n-1} S_\mu^{\ell-1}
-  \p{S+\mu S_\z} \p{-\mu+S_\z}^m S_\bmu^{n-2},
+  \p{S+\mu S_\z} \p{-1+\mu S_\z}^m S_\bmu^{n-2},
   \label{eq:san_z_mu} \allowdisplaybreaks \\
-  \sum_j \sigma_\z^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\bmu^{(j)}
-  &= -\mu S_\mu^\ell S_\z^m S_\bmu^{n+1}
-  + 2\mu \ell S_\mu^{\ell-1} \p{S+\mu S_\z}\p{-\mu+S_\z}^m S_\bmu^n,
+  \sum_j \sigma_\z^{(j)} S_\mu^\ell
+  \p{\mu S_\z}^n S_\bmu^n \sigma_\bmu^{(j)}
+  &= -\mu S_\mu^\ell \p{\mu S_\z}^m S_\bmu^{n+1}
+  + 2\mu \ell S_\mu^{\ell-1} \p{S+\mu S_\z}\p{-1+\mu S_\z}^m S_\bmu^n,
   \label{eq:san_z_bmu} \allowdisplaybreaks \\
-  \sum_j \sigma_\mu^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\mu^{(j)}
-  &= n S_\mu^{\ell+1} S_\z^m S_\bmu^{n-1}
-  - n\p{n-1} S_\mu^\ell \p{S+\mu S_\z} \p{-\mu+S_\z}^m S_\bmu^{n-2},
+  \sum_j \sigma_\mu^{(j)} S_\mu^\ell
+  \p{\mu S_\z}^n S_\bmu^n \sigma_\mu^{(j)}
+  &= n S_\mu^{\ell+1} \p{\mu S_\z}^m S_\bmu^{n-1}
+  - n\p{n-1} S_\mu^\ell \p{S+\mu S_\z} \p{-1+\mu S_\z}^m S_\bmu^{n-2},
   \label{eq:san_mu_mu} \allowdisplaybreaks \\
-  \sum_j \sigma_\mu^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\bmu^{(j)}
-  &= S_\mu^\ell \p{S+\mu S_\z}\p{-\mu+S_\z}^m S_\bmu^n,
+  \sum_j \sigma_\mu^{(j)} S_\mu^\ell
+  \p{\mu S_\z}^n S_\bmu^n \sigma_\bmu^{(j)}
+  &= S_\mu^\ell \p{S+\mu S_\z}\p{-1+\mu S_\z}^m S_\bmu^n,
   \label{eq:san_mu_bmu} \allowdisplaybreaks \\
-  \sum_j \sigma_\bmu^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\mu^{(j)}
-  &= S_\mu^\ell \p{S - \ell - n - \mu S_\z} \p{\mu + S_\z}^m S_\bmu^n
-  + \ell n \p{2S - \ell - n + 2}
-  S_\mu^{\ell-1} S_\z^m S_\bmu^{n-1} \notag \\
-  &\qquad + \ell n \p{\ell-1} \p{n-1} S_\mu^{\ell-2} \p{S + \mu S_\z}
-  \p{-\mu + S_\z}^m S_\bmu^{n-2}.
+  \sum_j \sigma_\bmu^{(j)} S_\mu^\ell
+  \p{\mu S_\z}^n S_\bmu^n \sigma_\mu^{(j)}
+  &= S_\mu^\ell \p{S - \ell - n - \mu S_\z}
+  \p{1+\mu S_\z}^m S_\bmu^n \notag \\
+  &\qquad + \ell n \p{2S - \ell - n + 2}
+  S_\mu^{\ell-1} \p{\mu S_\z}^m S_\bmu^{n-1} \notag \\
+  &\qquad + \ell n \p{\ell-1} \p{n-1} S_\mu^{\ell-2} \p{S+\mu S_\z}
+  \p{-1+\mu S_\z}^m S_\bmu^{n-2}.
   \label{eq:san_bmu_mu}
@@
 \begin{align}
-  \sp{S_\z, S_\mu^\ell S_\z^m S_\bmu^n}_+
-  = S_\z S_\mu^\ell S_\z^m S_\bmu^n + S_\mu^\ell S_\z^m S_\bmu^n S_\z
-  = S_\mu^\ell\p{\mu\ell + \mu n + 2 S_\z} S_\z^m S_\bmu^n,
+  \sp{S_\z, S_\mu^\ell \p{\mu S_\z}^n S_\bmu^n}_+
+  = S_\z S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n
+  + S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\z
+  = \mu S_\mu^\ell\p{\ell+n+2\mu S_\z} \p{\mu S_\z}^m S_\bmu^n,
 \end{align}
@@
 \begin{align}
-  \D\p{\sigma_\bmu; S_\mu^\ell S_\z^m S_\bmu^n}
-  = S_\mu^\ell \p{S+\mu S_\z}\p{-\mu+S_\z}^m S_\bmu^n
-  - S_\mu^\ell\sp{S + \f12\p{\ell+n} + \mu S_\z} S_\z^m S_\bmu^n,
+  \D\p{\sigma_\bmu; S_\mu^\ell \p{\mu S_\z}^n S_\bmu^n}
+  = S_\mu^\ell \p{S+\mu S_\z}\p{-1+\mu S_\z}^m S_\bmu^n
+  - S_\mu^\ell\sp{S + \f12\p{\ell+n} + \mu S_\z}
+  \p{\mu S_\z}^m S_\bmu^n,
   \label{eq:decay_diff}
@@
 \begin{align}
-  \D\p{\sigma_\mu; S_\mu^\ell S_\z^m S_\bmu^n}
-  &= S_\mu^\ell \p{S - \ell - n - \mu S_\z} \p{\mu + S_\z}^m S_\bmu^n
+  \D\p{\sigma_\mu; S_\mu^\ell \p{\mu S_\z}^n S_\bmu^n}
+  &= S_\mu^\ell \p{S - \ell - n - \mu S_\z} \p{1+\mu S_\z}^m S_\bmu^n
   - S_\mu^\ell\sp{S - \f12\p{\ell+n} - \mu S_\z}
-  S_\z^m S_\bmu^n \notag \\
+  \p{\mu S_\z}^m S_\bmu^n \notag \\
   &\qquad + \ell n \p{2S - \ell - n + 2}
-  S_\mu^{\ell-1} S_\z^m S_\bmu^{n-1} \notag \\
+  S_\mu^{\ell-1} \p{\mu S_\z}^m S_\bmu^{n-1} \notag \\
   &\qquad + \ell n \p{\ell-1} \p{n-1} S_\mu^{\ell-2} \p{S + \mu S_\z}
-  \p{-\mu + S_\z}^m S_\bmu^{n-2}.
+  \p{-1+\mu S_\z}^m S_\bmu^{n-2}.
   \label{eq:decay_same}
@@
 Decoherence via jump operators $\sigma_\bmu^{(j)}$ only couples
-operators $S_\mu^\ell S_\z^m S_\bmu^n$ to operators
-$S_\mu^\ell S_\z^{m'} S_\bmu^n$ with $m'\le m$.  Decoherence via jump
-operators $\sigma_\mu^{(j)}$, meanwhile, makes operators
-$S_\mu^\ell S_\z^m S_\bmu^n$ ``grow'' in $S_\z$ through the last term
-in \eqref{eq:decay_same} with $k=m$, although the sum $\ell+m+n$ does
-not grow.
+operators $S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$ to operators
+$S_\mu^\ell \p{\mu S_\z}^{m'} S_\bmu^n$ with $m'\le m$.  Decoherence
+via jump operators $\sigma_\mu^{(j)}$, meanwhile, makes operators
+$S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$ ``grow'' in $m$ through the last
+term in \eqref{eq:decay_same} with $k=m$, although the sum $\ell+m+n$
+does not grow.
 
@@
 \begin{align}
-  \D\p{\sigma_\z/\sqrt2; S_\mu^\ell S_\z^m S_\bmu^n}
-  = -\p{\ell+n} S_\mu^\ell S_\z^m S_\bmu^n
+  \D\p{\sigma_\z/\sqrt2; S_\mu^\ell \p{\mu S_\z}^n S_\bmu^n}
+  = -\p{\ell+n} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n
   + 2 \ell n S_\mu^{\ell-1} \p{S + \mu S_\z}
-  \p{-\mu + S_\z}^m S_\bmu^{n-1}.
+  \p{-1 + \mu S_\z}^m S_\bmu^{n-1}.
 \end{align}
 Decoherence via single-spin dephasing makes operators
-$S_\mu^\ell S_\z^m S_\bmu^n$ ``grow'' in $S_\z$, although the sum
+$S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$ ``grow'' in $m$, although the sum
 $\ell+m+n$ does not grow.
@@
 \begin{align}
-  \sp{S_\mu, S_\mu^\ell S_\z^m S_\bmu^n}_+
-  = S_\mu^{\ell+1} \sp{S_\z^m+\p{\mu+S_\z}^m} S_\bmu^n
-  - n S_\mu^\ell \p{n-1+2\mu S_\z} S_\z^m S_\bmu^{n-1}.
+  \sp{S_\mu, S_\mu^\ell \p{\mu S_\z}^n S_\bmu^n}_+
+  = S_\mu^{\ell+1} \sp{S_\z^m+\p{1+\mu S_\z}^m} S_\bmu^n
+  - n S_\mu^\ell \p{n-1+2\mu S_\z} \p{\mu S_\z}^m S_\bmu^{n-1}.
   \label{eq:S_mu_acomm}
@@
 \begin{align}
-  \sum_j \sigma_\z^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\mu^{(j)}
-  - \f{\mu}{2} \sp{S_\mu, S_\mu^\ell S_\z^m S_\bmu^n}_+
+  \sum_j \sigma_\z^{(j)} S_\mu^\ell
+  \p{\mu S_\z}^n S_\bmu^n \sigma_\mu^{(j)}
+  - \f{\mu}{2} \sp{S_\mu, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
   = K_{\ell mn;\mu} + L_{\ell mn;\mu}
@@
 \begin{align}
-  \sum_j \sigma_\mu^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\z^{(j)}
-  + \f{\mu}{2} \sp{S_\mu, S_\mu^\ell S_\z^m S_\bmu^n}_+
+  \sum_j \sigma_\mu^{(j)} S_\mu^\ell
+  \p{\mu S_\z}^n S_\bmu^n \sigma_\z^{(j)}
+  + \f{\mu}{2} \sp{S_\mu, S_\mu^\ell \p{\mu S_\z}^n S_\bmu^n}_+
   = K_{\ell mn;\mu} + M_{\ell mn;\mu}
@@
   &\equiv \f{\mu}{2} S_\mu^{\ell+1}
-  \sp{\p{\mu+S_\z}^m-S_\z^m} S_\bmu^n, \\
+  \sp{\p{1+\mu S_\z}^m-\p{\mu S_\z}^m} S_\bmu^n, \\
   L_{\ell mn;\mu}
   &\equiv \mu n S_\mu^\ell \sp{-2S+2\ell+\f32\p{n-1}+\mu S_\z}
-  S_\z^m S_\bmu^{n-1} \notag \\
+  \p{\mu S_\z}^m S_\bmu^{n-1} \notag \\
   &\qquad - 2\mu\ell n\p{n-1} S_\mu^{\ell-1}
-  \p{S+\mu S_\z} \p{-\mu+S_\z}^m S_\bmu^{n-2}, \\
+  \p{S+\mu S_\z} \p{-1+\mu S_\z}^m S_\bmu^{n-2}, \\
   M_{\ell mn;\mu}
-  &\equiv \mu n S_\mu^\ell \sp{2\p{S+\mu S_\z}\p{-\mu+S_\z}^m
-    - \p{\f12\sp{n-1}+\mu S_\z} S_\z^m} S_\bmu^{n-1}.
+  &\equiv \mu n S_\mu^\ell \sp{2\p{S+\mu S_\z}\p{-1+\mu S_\z}^m
+    - \p{\f12\sp{n-1}+\mu S_\z}\p{\mu S_\z}^m} S_\bmu^{n-1}.
 \end{align}
@@
   P_{\ell mn;\mu}
-  \equiv \sum_j
-  \sigma_\mu^{(j)} S_\mu^\ell S_\z^m S_\bmu^n \sigma_\mu^{(j)}
-  = n S_\mu^{\ell+1} S_\z^m S_\bmu^{n-1}
-  - n\p{n-1} S_\mu^\ell \p{S+\mu S_\z} \p{-\mu+S_\z}^m S_\bmu^{n-2},
+  &\equiv \sum_j \sigma_\mu^{(j)} S_\mu^\ell
+  \p{\mu S_\z}^m S_\bmu^n \sigma_\mu^{(j)} \\
+  &= n S_\mu^{\ell+1} \p{\mu S_\z}^m S_\bmu^{n-1}
+  - n\p{n-1} S_\mu^\ell \p{S+\mu S_\z} \p{-1+\mu S_\z}^m S_\bmu^{n-2},
 \end{align}
@@
 \begin{align}
-  \D\p{\gamma; S_\mu^\ell S_\z^m S_\bmu^n}
+  \D\p{\gamma; S_\mu^\ell \p{\mu S_\z}^n S_\bmu^n}
   = \sum_{X\in\set{\z,+,-}} \abs{\gamma_X}^2
-  \D\p{\sigma_X; S_\mu^\ell S_\z^m S_\bmu^n}
+  \D\p{\sigma_X; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
   + Q_{\ell mn;\mu}^{(\gamma)} + \sp{Q_{nm\ell;\mu}^{(\gamma)}}^\dag.
 \end{align}
-Note that the sum $\ell+m+n$ does not ever grow under this form of
+Note that the sum $\ell+m+n$ for operators
+$S_\mu^\ell \p{\mu S_\z}^n S_\bmu^n$ does not grow under this type of
 decoherence.
@@
 \begin{align}
-  S_\mu S_\mu^\ell S_\z^m S_\bmu^n S_\bmu
-  &= S_\mu^{\ell+1} S_\z^m S_\bmu^{n+1},
+  S_\mu S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\bmu
+  &= S_\mu^{\ell+1} \p{\mu S_\z}^m S_\bmu^{n+1},
   \allowdisplaybreaks \\
-  S_\mu S_\mu^\ell S_\z^m S_\bmu^n S_\z
-  &= S_\mu^{\ell+1} \p{\mu n+S_\z} S_\z^m S_\bmu^n,
+  S_\mu S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\z
+  &= \mu S_\mu^{\ell+1} \p{n+\mu S_\z} \p{\mu S_\z}^m S_\bmu^n,
   \allowdisplaybreaks \\
-  S_\z S_\mu^\ell S_\z^m S_\bmu^n S_\z
-  &= S_\mu^\ell \sp{\ell n + \mu\p{\ell+n} S_\z + S_\z^2}
-  S_\z^m S_\bmu^n.
+  S_\z S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\z
+  &= S_\mu^\ell \sp{\ell n + \p{\ell+n} \mu S_\z + \p{\mu S_\z}^2}
+  \p{\mu S_\z}^m S_\bmu^n.
 \end{align}
@@
 \begin{align}
-  S_\mu^\ell S_\z^m S_\bmu^n S_\mu
-  &= S_\mu^{\ell+1} \p{\mu+S_\z}^m S_\bmu^n
-  - n S_\mu^\ell \p{n-1+2\mu S_\z} S_\z^m S_\bmu^{n-1},
+  S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\mu
+  &= S_\mu^{\ell+1} \p{1+\mu S_\z}^m S_\bmu^n
+  - n S_\mu^\ell \p{n-1+2\mu S_\z} \p{\mu S_\z}^m S_\bmu^{n-1},
 \end{align}
@@
 \begin{align}
-  S_\mu S_\mu^\ell S_\z^m S_\bmu^n S_\mu
-  &= S_\mu^{\ell+2} \p{\mu+S_\z}^m S_\bmu^n
-  - n S_\mu^{\ell+1} \p{n-1+2\mu S_\z} S_\z^m S_\bmu^{n-1},
+  S_\mu S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\mu
+  &= S_\mu^{\ell+2} \p{1+\mu S_\z}^m S_\bmu^n
+  - n S_\mu^{\ell+1} \p{n-1+2\mu S_\z} \p{\mu S_\z}^m S_\bmu^{n-1},
   \allowdisplaybreaks \\
-  S_\z S_\mu^\ell S_\z^m S_\bmu^n S_\mu
-  &= S_\mu^{\ell+1} \sp{\mu\p{\ell+1}+S_\z} \p{\mu+S_\z}^m S_\bmu^n
+  S_\z S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\mu
+  &= \mu S_\mu^{\ell+1} \p{\ell+1+\mu S_\z} \p{1+\mu S_\z}^m S_\bmu^n
   \notag \\
-  &\qquad - n S_\mu^\ell
-  \sp{\mu\ell\p{n-1} + \p{2\ell+n-1}S_\z + 2\mu S_\z^2}
-  S_\z^m S_\bmu^{n-1}.
+  &\qquad - \mu n S_\mu^\ell
+  \sp{\ell\p{n-1} + \p{2\ell+n-1}\mu S_\z + 2\p{\mu S_\z}^2}
+  \p{\mu S_\z}^m S_\bmu^{n-1}.
 \end{align}
@@
 \begin{align}
-  S_\bmu S_\mu^\ell S_\z^m S_\bmu^n S_\mu
+  S_\bmu S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\mu
   &= \sp{S_\mu^\ell S_\bmu - \ell S_\mu^{\ell-1} \p{\ell-1+2\mu S_\z}}
-  S_\z^m
-  \sp{S_\mu S_\bmu^n - n \p{n-1+2\mu S_\z} S_\bmu^{n-1}} \\
-  &= S_\mu^\ell S_\bmu S_\z^m S_\mu S_\bmu^n
-  - S_\mu^\ell \sp{\ell\p{\ell+1} + n\p{n+1}+2\mu\p{\ell+n} S_\z}
-  \p{\mu+S_\z}^m S_\bmu^n \notag \\
+  \p{\mu S_\z}^m
+  \sp{S_\mu S_\bmu^n - n \p{n-1+2\mu S_\z} S_\bmu^{n-1}} \notag \\
+  &= S_\mu^\ell S_\bmu \p{\mu S_\z}^m S_\mu S_\bmu^n \notag \\
+  &\qquad - S_\mu^\ell
+  \sp{\ell\p{\ell+1} + n\p{n+1}+2\p{\ell+n}\mu S_\z}
+  \p{1+\mu S_\z}^m S_\bmu^n \notag \\
   &\qquad + \ell n S_\mu^{\ell-1}
-  \sp{\p{\ell-1}\p{n-1}+2\mu\p{\ell+n-2} S_\z + 4 S_\z^2}
-  S_\z^m S_\bmu^{n-1},
+  \sp{\p{\ell-1}\p{n-1}+2\p{\ell+n-2}\mu S_\z + 4\p{\mu S_\z}^2}
+  \p{\mu S_\z}^m S_\bmu^{n-1},
 \end{align}
 where
-\begin{align}
-  S_\bmu S_\z^m S_\mu
-  = S_\bmu S_\mu \p{\mu+S_\z}^m
-  = \p{S_\mu S_\bmu - 2\mu S_\z} \p{\mu+S_\z}^m
-  = S_\mu \p{2\mu + S_\z}^m S_\bmu - 2\mu S_\z \p{\mu + S_\z}^m,
-\end{align}
+\begin{multline}
+  S_\bmu \p{\mu S_\z}^m S_\mu
+  = S_\bmu S_\mu \p{1+\mu S_\z}^m
+  = \p{S_\mu S_\bmu - 2\mu S_\z} \p{1+\mu S_\z}^m \\
+  = S_\mu \p{2+\mu S_\z}^m S_\bmu - 2\mu S_\z \p{1+\mu S_\z}^m,
+\end{multline}
 so
 \begin{align}
-  S_\bmu S_\mu^\ell S_\z^m S_\bmu^n S_\mu
-  &= S_\mu^{\ell+1} \p{2\mu + S_\z}^m S_\bmu^{n+1} \notag \\
+  S_\bmu S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\mu
+  &= S_\mu^{\ell+1} \p{2+\mu S_\z}^m S_\bmu^{n+1} \notag \\
   &\qquad - S_\mu^\ell
-  \sp{\ell\p{\ell+1} + n\p{n+1}+2\mu\p{\ell+n+1} S_\z}
-  \p{\mu+S_\z}^m S_\bmu^n \notag \\
+  \sp{\ell\p{\ell+1} + n\p{n+1}+2\p{\ell+n+1}\mu S_\z}
+  \p{1+\mu S_\z}^m S_\bmu^n \notag \\
   &\qquad + \ell n S_\mu^{\ell-1}
-  \sp{\p{\ell-1}\p{n-1}+2\mu\p{\ell+n-2} S_\z + 4 S_\z^2}
-  S_\z^m S_\bmu^{n-1}.
+  \sp{\p{\ell-1}\p{n-1}+2\p{\ell+n-2}\mu S_\z + 4\p{\mu S_\z}^2}
+  \p{\mu S_\z}^m S_\bmu^{n-1}.
 \end{align}
@@
 \begin{align}
-  \D\p{S_\bmu; S_\mu^\ell S_\z^m S_\bmu^n}
-  &= -S_\mu^{\ell+1} \sp{\p{\mu+S_\z}^m - S_\z^m} S_\bmu^{n+1} \notag \\
+  \D\p{S_\bmu; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
+  &= -S_\mu^{\ell+1} \sp{\p{1+\mu S_\z}^m - \p{\mu S_\z}^m}
+  S_\bmu^{n+1} \notag \\
   &\qquad + \f12 S_\mu^\ell \sp{\ell\p{\ell-1} + n\p{n-1}
-    + 2\mu\p{\ell+n} S_\z} S_\z^m S_\bmu^n,
+    + 2\p{\ell+n}\mu S_\z} \p{\mu S_\z}^m S_\bmu^n,
 \end{align}
@@
 \begin{align}
-  \D\p{S_\mu; S_\mu^\ell S_\z^m S_\bmu^n}
-  &= S_\mu^{\ell+1} \sp{\p{2\mu + S_\z}^m-\p{\mu + S_\z}^m}
+  \D\p{S_\mu; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
+  &= S_\mu^{\ell+1} \sp{\p{2+\mu S_\z}^m-\p{1+\mu S_\z}^m}
   S_\bmu^{n+1} \notag \\
   &\qquad - S_\mu^\ell
-  \sp{\ell\p{\ell+1} + n\p{n+1}+2\mu\p{\ell+n+1} S_\z}
-  \p{\mu+S_\z}^m S_\bmu^n \notag \\
+  \sp{\ell\p{\ell+1} + n\p{n+1}+2\p{\ell+n+1}\mu S_\z}
+  \p{1+\mu S_\z}^m S_\bmu^n \notag \\
   &\qquad + \f12 S_\mu^\ell
-  \sp{\ell\p{\ell+1} + n\p{n+1}+2\mu\p{\ell+n+2} S_\z}
-  S_\z^m S_\bmu^n \notag \\
+  \sp{\ell\p{\ell+1} + n\p{n+1}+2\p{\ell+n+2}\mu S_\z}
+  \p{\mu S_\z}^m S_\bmu^n \notag \\
   &\qquad + \ell n S_\mu^{\ell-1}
-  \sp{\p{\ell-1}\p{n-1}+2\mu\p{\ell+n-2} S_\z + 4 S_\z^2}
-  S_\z^m S_\bmu^{n-1}.
+  \sp{\p{\ell-1}\p{n-1}+2\p{\ell+n-2}\mu S_\z + 4\p{\mu S_\z}^2}
+  \p{\mu S_\z}^m S_\bmu^{n-1}.
 \end{align}
@@
 \begin{align}
-  \D\p{S_\z; S_\mu^\ell S_\z^m S_\bmu^n}
-  = -\f12 \p{\ell-n}^2 S_\mu^\ell S_\z^m S_\bmu^n.
+  \D\p{S_\z; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
+  = -\f12 \p{\ell-n}^2 S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n.
 \end{align}
@@
 \begin{align}
-  \sp{S_\mu^2, S_\mu^\ell S_\z^m S_\bmu^n}_+
-  &= S_\mu^{\ell+2} \sp{S_\z^m+\p{2\mu+S_\z}^m} S_\bmu^n
-  - 2n S_\mu^{\ell+1} \p{n+2\mu S_\z} \p{\mu+S_\z}^m S_\bmu^{n-1}
+  \sp{S_\mu^2, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
+  &= S_\mu^{\ell+2} \sp{\p{2+\mu S_\z}^m+\p{\mu S_\z}^m} S_\bmu^n
+  - 2n S_\mu^{\ell+1} \p{n+2\mu S_\z} \p{1+\mu S_\z}^m S_\bmu^{n-1}
   \notag \\
   &\qquad + n\p{n-1} S_\mu^\ell \sp{\p{n-1}\p{n-2}
-    + 2\mu\p{2n-3} S_\z + 4 S_\z^2} S_\z^m S_\bmu^{n-2},
+    + 2\p{2n-3}\mu S_\z + 4\p{\mu S_\z}^2} \notag \\
+  &\qquad\qquad\qquad\qquad \times \p{\mu S_\z}^m S_\bmu^{n-2},
   \allowdisplaybreaks \\
-  \sp{S_\z S_\mu, S_\mu^\ell S_\z^m S_\bmu^n}_+
-  &= S_\mu^{\ell+1} \sp{\p{\mu\sp{\ell+1} + S_\z} S_\z^m
-    + \p{\mu\sp{n+1} + S_\z} \p{\mu+S_\z}^m } S_\bmu^n \notag \\
-  &\qquad - n S_\mu^\ell \sp{\mu n \p{n-1}
-    + \p{3n-1} S_\z + 2\mu S_\z^2} S_\z^m S_\bmu^{n-1},
+  \sp{S_\z S_\mu, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
+  &= \mu S_\mu^{\ell+1} \sp{\p{\ell+1+\mu S_\z}\p{\mu S_\z}^m
+    + \p{n+1+\mu S_\z} \p{1+\mu S_\z}^m } S_\bmu^n \notag \\
+  &\qquad - \mu n S_\mu^\ell \sp{n \p{n-1}
+    + \p{3n-1}\mu S_\z + 2\p{\mu S_\z}^2} \p{\mu S_\z}^m S_\bmu^{n-1},
   \allowdisplaybreaks \\
-  \sp{S_\mu S_\z, S_\mu^\ell S_\z^m S_\bmu^n}_+
-  &= S_\mu^{\ell+1} \sp{\p{\mu\ell+S_\z} S_\z^m
-    + \p{\mu n+S_\z} \p{\mu+S_\z}^m} S_\bmu^n \notag \\
-  &\qquad - n S_\mu^\ell \sp{\mu\p{n-1}^2
-    + 3\p{n-1} S_\z + 2\mu S_\z^2} S_\z^m S_\bmu^{n-1}.
+  \sp{S_\mu S_\z, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
+  &= \mu S_\mu^{\ell+1} \sp{\p{\ell+\mu S_\z}\p{\mu S_\z}^m
+    + \p{n+\mu S_\z} \p{1+\mu S_\z}^m} S_\bmu^n \notag \\
+  &\qquad - \mu n S_\mu^\ell \sp{\p{n-1}^2
+    + 3\p{n-1}\mu S_\z + 2\p{\mu S_\z}^2} \p{\mu S_\z}^m S_\bmu^{n-1}.
 \end{align}
@@
     + \p{\ell+\f12} \Gamma_{\z,\mu}^{(-)}}
-  S_\mu^{\ell+1} \p{\mu+S_\z}^m S_\bmu^n \notag \\
+  S_\mu^{\ell+1} \p{1+\mu S_\z}^m S_\bmu^n \notag \\
   &\qquad -\mu \sp{\p{\ell-n+\f12} \Gamma_{\z,\mu}^{(+)}
     + \p{n+\f12} \Gamma_{\z,\mu}^{(-)}}
-  S_\mu^{\ell+1} S_\z^m S_\bmu^n \notag \\
-  &\qquad + \Gamma_{\z,\mu}^{(-)}
-  S_\mu^{\ell+1} S_\z \sp{\p{\mu+S_\z}^m - S_\z^m} S_\bmu^n,
+  S_\mu^{\ell+1} \p{\mu S_\z}^m S_\bmu^n \notag \\
+  &\qquad + \mu \Gamma_{\z,\mu}^{(-)}
+  S_\mu^{\ell+1} \p{\mu S_\z}
+  \sp{\p{1+\mu S_\z}^m - \p{\mu S_\z}^m} S_\bmu^n,
   \allowdisplaybreaks \\
@@
     + \p{\ell-\f12} \Gamma_{\z,\mu}^{(-)}}
-  S_\mu^\ell S_\z^m S_\bmu^{n-1} \notag \\
-  &\qquad - 2n \sp{\p{\ell-n+\f12} \Gamma_{\z,\mu}^{(+)}
+  S_\mu^\ell \p{\mu S_\z}^m S_\bmu^{n-1} \notag \\
+  &\qquad - 2\mu n \sp{\p{\ell-n+\f12} \Gamma_{\z,\mu}^{(+)}
     + \p{\ell+\f12n-1} \Gamma_{\z,\mu}^{(-)}}
-  S_\mu^\ell S_\z^{m+1} S_\bmu^{n-1} \notag \\
+  S_\mu^\ell \p{\mu S_\z}^{m+1} S_\bmu^{n-1} \notag \\
   &\qquad - 2\mu n \Gamma_{\z,\mu}^{(-)}
-  S_\mu^\ell S_\z^{m+2} S_\bmu^{n-1},
+  S_\mu^\ell \p{\mu S_\z}^{m+2} S_\bmu^{n-1},
   \allowdisplaybreaks \\
@@
   &\equiv -\f12 S_\mu^{\ell+2}
-  \sp{\p{2\mu+S_\z}^m - 2\p{\mu+S_\z}^m + S_\z^m} S_\bmu^n \notag \\
-  &\qquad + n S_\mu^{\ell+1} \sp{\p{n+2\mu S_\z} \p{\mu+S_\z}^m
-    - \p{n-1+2\mu S_\z} S_\z^m} S_\bmu^{n-1} \notag \\
+  \sp{\p{2+\mu S_\z}^m - 2\p{1+\mu S_\z}^m + \p{\mu S_\z}^m}
+  S_\bmu^n \notag \\
+  &\qquad + n S_\mu^{\ell+1} \sp{\p{n+2\mu S_\z} \p{1+\mu S_\z}^m
+    - \p{n-1+2\mu S_\z} \p{\mu S_\z}^m}
+  S_\bmu^{n-1} \notag \\
   &\qquad -n\p{n-1} S_\mu^\ell
-  \sp{\f12\p{n-1}\p{n-2} + \mu\p{2n-3} S_\z + 2 S_\z^2}
-  S_\z^m S_\bmu^{n-2},
+  \sp{\f12\p{n-1}\p{n-2} + \p{2n-3}\mu S_\z + 2\p{\mu S_\z}^2}
+  \p{\mu S_\z}^m S_\bmu^{n-2},
   \allowdisplaybreaks \\
@@
 \begin{align}
-  \D\p{\Gamma; S_\mu^\ell S_\z^m S_\bmu^n}
+  \D\p{\Gamma; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
   = \sum_{X\in\set{\z,+,-}} \abs{\Gamma_X}^2
-  \D\p{S_X; S_\mu^\ell S_\z^m S_\bmu^n}
+  \D\p{S_X; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
   + \tilde Q_{\ell mn;\mu}^{(\Gamma)}
@@
 \end{align}
-Note that the sum $\ell+m+n$ grows by one if $\Gamma_\mu\ne0$ or
-$\Gamma_\bmu\ne0$, and does not grow otherwise.
+Note that the sum $\ell+m+n$ for operators
+$S_\mu^\ell \p{\mu S_\z}^n S_\bmu^n$ grows by one if $\Gamma_\mu\ne0$
+or $\Gamma_\bmu\ne0$, and does not grow otherwise.
 
@@
 \end{align}
-using which we can compute, for $\mu,\nu\in\set{+1,-1}$ and
+using which we can compute, for $\mu,\nu\in\set{+1,-1}$,
 $\bmu\equiv-\mu$,
 \begin{align}
-  \bk{\nu\Z|S_\mu^\ell S_\z^m S_\bmu^n|\nu\Z}
-  = \delta_{\ell n} \nu^m \times
+  \bk{\nu\Z|S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n|\nu\Z}
+  = \delta_{\ell n} \times
   \begin{cases}
      \p{S-n}^m \f{N! n!}{\p{N-n}!} & \mu = \nu, \\
-     \delta_{n,0} S^m & \mu \ne \nu,
+     \delta_{n,0} \p{-S}^m & \mu \ne \nu,
   \end{cases}
@@
 \begin{align}
-  \bk{\X|S_\mu^\ell S_\z^m S_\bmu^n|\X}
-  &=  2^{-N} \sum_{k=-S+\delta_{\mu,-1}\max\set{\ell,n}}
+  \bk{\X|S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n|\X}
+  &=  \f{\mu^m}{2^N} \sum_{k=-S+\delta_{\mu,-1}\max\set{\ell,n}}
   ^{S-\delta_{\mu,1}\max\set{\ell,n}} k^m
@@
   \prod_{q=0}^{n-1} g_\bmu\p{S,k+\mu\sp{n-q}} \\
-  &= \f{N!}{2^N} \sum_{k=-S+\delta_{\mu,-1}\max\set{\ell,n}}
+  &= \mu^m \f{N!}{2^N} \sum_{k=-S+\delta_{\mu,-1}\max\set{\ell,n}}
   ^{S-\delta_{\mu,1}\max\set{\ell,n}} k^m
   \f{\p{S-\mu k}!}{\p{S+\mu k}!\p{S-\mu k-\ell}!\p{S-\mu k-n}!} \\
-  &= \p{-\mu}^m \f{N!}{2^N} \sum_{k=-S+\max\set{\ell,n}}^S k^m
+  &= \p{-1}^m \f{N!}{2^N} \sum_{k=-S+\max\set{\ell,n}}^S k^m
   \f{\p{S+k}!}{\p{S-k}!\p{S+k-\ell}!\p{S+k-n}!}.
@@
 \begin{align}
-  \bk{\phi|S_\mu^\ell S_\z^m S_\bmu^n|\phi}
-  = e^{i\mu\p{\ell-n}\phi} \bk{\X|S_\mu^\ell S_\z^m S_\bmu^n|\X},
+  \bk{\phi|S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n|\phi}
+  = e^{i\mu\p{\ell-n}\phi} \bk{\X|S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n|\X},
 \end{align}
@@
 \begin{align}
-  \bk{\X|S_\mu^\ell S_\z^m S_\bmu^n|\X}
-  = \f{\mu^m\p{N!}^2}{2^N\p{N-\ell}!\p{N-n}!}
+  \bk{\X|S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n|\X}
+  = \f{\p{N!}^2}{2^N\p{N-\ell}!\p{N-n}!}
   \sp{\p{-S}^m \alpha + \f{\p{N-\ell}\p{N-n}}{N} \beta},

### 2018-09-14 soc_squeezing/scripts/correlator_methods.py

@@ -46,8 +46,8 @@ def stirling(nn, kk, stirlings = {}):
 def zeta(mm, nn, pp, qq, zetas = {}):
     try: return zetas[(mm,nn,pp,qq)]
     except:
-        prefactor = (-1)**(pp-qq) * 2**qq
-        sum_ss = sum([ binom(ss,qq) * stirling(pp,ss) * (mm+nn-2*pp)**(ss-qq)
+        prefactor = (-1)**pp * 2**qq
+        sum_ss = sum([ stirling(pp,ss) * binom(ss,qq) * (mm+nn-2*pp)**(ss-qq)
                        for ss in range(qq,pp+1) ])
         zetas[(mm,nn,pp,qq)] = prefactor * sum_ss
         zetas[(nn,mm,pp,qq)] = prefactor * sum_ss
@@ -74,7 +74,7 @@ def op_val_X(op, SS):
     ln_prefactor = ln_factorial(2*SS) - 2*SS*np.log(2)
     k_vals = np.arange(-SS+max(ll,nn),SS+0.5)
     terms = k_vals**mm * np.exp(ln_factors_X(SS,k_vals,ll,nn) + ln_prefactor)
-    return terms.sum()
+    return (-1)**mm * terms.sum()
 
 # correlator < Z | S_+^ll S_\z^mm S_-^nn | Z >
 def op_val_Z_p(op, SS):
@@ -90,31 +90,43 @@ def op_val_Z_p(op, SS):
 def op_val_Z_m(op, SS):
     ll, mm, nn = op
     if ll != 0 or nn != 0: return 0
-    return SS**mm
+    return (-SS)**mm
 
-# correlator ln < X | S_+^ll S_\z^mm S_-^nn | X >
+# correlator ln | < X | S_+^ll S_\z^mm S_-^nn | X > |
 def op_ln_val_X(op, SS):
     ll, mm, nn = op
     if ll == 0 and nn == 0 and mm % 2 == 1: return None
     if max(ll,nn) > 2*SS: return None
 
-    offset = SS % 1
-    k_vals_p = np.arange(1-offset,SS+offset)
-    k_vals_n = np.arange(1-offset,SS-max(ll,nn)+offset)
-
-    sign_n = ( 1 if mm % 2 == 0 else -1 )
     ln_prefactor = ln_factorial(2*SS) - 2*SS*np.log(2)
     ln_factors = lambda kk : ln_factors_X(SS,kk,ll,nn)
+
[…]

### 2018-09-14 soc_squeezing/scripts/test_commutator.py

@@ -1,8 +1,6 @@
 #!/usr/bin/env python3
 
-import sys, scipy
 import numpy as np
-import matplotlib.pyplot as plt
 
 import itertools
 from scipy.special import factorial, binom
@@ -15,72 +13,76 @@ from squeezing_methods import val
 N = 50
 S = N/2
 
-def stirling(n,k):
-    return float(sympy_stirling(n, k, kind = 1, signed = False))
+def stirling(nn,kk):
+    return float(sympy_stirling(nn, kk, kind = 1, signed = False))
 
 II = sparse.identity(N+1)
 ZZ = sparse.dok_matrix((N+1,N+1))
 
-Sz = spin_op_z_dicke(N)
-Sp = spin_op_p_dicke(N)
-Sm = spin_op_m_dicke(N)
+S_z = spin_op_z_dicke(N)
+S_p = spin_op_p_dicke(N)
+S_m = spin_op_m_dicke(N)
 
-Sz_r = [ II ]
-Sp_r = [ II ]
-Sm_r = [ II ]
+S_z_r = [ II ]
+S_p_r = [ II ]
+S_m_r = [ II ]
 for n in range(N):
-    Sz_r.append(Sz_r[n].dot(Sz))
-    Sp_r.append(Sp_r[n].dot(Sp))
-    Sm_r.append(Sm_r[n].dot(Sm))
+    S_z_r.append(S_z_r[n].dot(S_z))
+    S_p_r.append(S_p_r[n].dot(S_p))
+    S_m_r.append(S_m_r[n].dot(S_m))
 
+
+def S_op(mu):
+    if mu == 1: return S_p
+    else: return S_m
 
 def S_ops(mu):
-    if mu == 1: return Sp_r, Sm_r
-    else: return Sm_r, Sp_r
+    return S_op(mu), S_op(-mu)
 
-def eps(m,n,p,l):
-    num = np.sum([ (-1)**(p-qq) * stirling(p,qq) * binom(qq,l) * (m-n)**(qq-l)
-                   for qq in range(l,p+1) ])
-    return 2**l * num
+def res(X,Y):
+    diff = X - Y
+    return abs(diff).max() / abs(X).max()
 
-def xi(m,n,p,q):
-    return sum([ (-1)**(ll-q) * eps(m,n,p,ll) * binom(ll,q) * (m-p)**(ll-q)
-                 for ll in range(q,p+1) ])
+def eps(mm,nn,pp,ll):
[…]

### 2018-09-14 soc_squeezing/scripts/test_decoherence.py

@@ -66,7 +66,7 @@ for l, m, n in itertools.product(range(N), repeat = 3):
 
         S_mu, S_nu, s_mu_j, s_nu_j = ops(mu)
 
-        op = S_mu**l * S_z**m * S_nu**n
+        op = S_mu**l * (mu*S_z)**m * S_nu**n
         g_ops = [ g_z * s_z_j[jj] + g_p * s_mu_j[jj] + g_m * s_nu_j[jj] for jj in range(N) ]
         G_op = g_z * S_z + g_p * S_mu + g_m * S_nu
 
@@ -74,58 +74,50 @@ for l, m, n in itertools.product(range(N), repeat = 3):
         A = D_full([G_op], op)
 
         ##############################
-
-        d_p = S_mu**l * ( (S-l-n-mu*S_z) * sum( mu**(m-k) * binom(m,k) * S_z**k
-                                                for k in range(m) )
-                          - 1/2*(l+n) * S_z**m ) * S_nu**n
+        d_p = S_mu**l * (S-l-n-mu*S_z) * (1 + mu*S_z)**m * S_nu**n
+        d_p -= S_mu**l * ( S - (l+n)/2 - mu*S_z ) * (mu*S_z)**m * S_nu**n
         if l >= 1 and n >= 1:
-            d_p += l*n*(2*S-l-n+2) * S_mu**(l-1) * S_z**m * S_nu**(n-1)
+            d_p += l*n*(2*S-l-n+2) * S_mu**(l-1) * (mu*S_z)**m * S_nu**(n-1)
         if l >= 2 and n >= 2:
-            d_p += ( l*n*(l-1)*(n-1) * S_mu**(l-2) * (S+mu*S_z)
-                     * sum( (-mu)**(m-k) * binom(m,k) * S_z**k for k in range(m+1) )
-                     * S_nu**(n-2) )
+            d_p += l*n*(l-1)*(n-1) * S_mu**(l-2) * (S+mu*S_z) * (-1+mu*S_z)**m * S_nu**(n-2)
 
         ##############################
 
-        d_m = S_mu**l * ( (S+mu*S_z) * sum( (-mu)**(m-k) * binom(m,k) * S_z**k
-                                            for k in range(m) )
-                          - 1/2 * (l+n) * S_z**m ) * S_nu**n
[…]

### 2018-09-14 soc_squeezing/scripts/correlator_methods.py
changed convention for sum limits in expectation value w.r.t. X

@@ -60,10 +60,10 @@ def zeta(mm, nn, pp, qq, zetas = {}):
 
 # natural logarithm of factors which appear in the expetation value for |X>
 def ln_factors_X(SS, kk, ll, nn):
-    return ( ln_factorial(SS + kk)
-             - ln_factorial(SS - kk)
-             - ln_factorial(SS + kk - ll)
-             - ln_factorial(SS + kk - nn) )
+    return ( ln_factorial(SS - kk)
+             - ln_factorial(SS + kk)
+             - ln_factorial(SS - kk - ll)
+             - ln_factorial(SS - kk - nn) )
 ln_factors_X = np.vectorize(ln_factors_X)
 
 # correlator < X | S_+^ll S_\z^mm S_-^nn | X >
@@ -72,9 +72,9 @@ def op_val_X(op, SS):
     if ll == 0 and nn == 0 and mm % 2 == 1: return 0
     if max(ll,nn) > 2*SS: return 0
     ln_prefactor = ln_factorial(2*SS) - 2*SS*np.log(2)
-    k_vals = np.arange(-SS+max(ll,nn),SS+0.5)
+    k_vals = np.arange(-SS,SS-max(ll,nn)+0.5)
     terms = k_vals**mm * np.exp(ln_factors_X(SS,k_vals,ll,nn) + ln_prefactor)
-    return (-1)**mm * terms.sum()
+    return terms.sum()
 
 # correlator < Z | S_+^ll S_\z^mm S_-^nn | Z >
 def op_val_Z_p(op, SS):
@@ -102,29 +102,29 @@ def op_ln_val_X(op, SS):
     ln_factors = lambda kk : ln_factors_X(SS,kk,ll,nn)
 
     if mm == 0: # this is a special case due to the contribution from the k = 0 term
-        k_vals = np.arange(-SS+max(ll,nn),SS+0.5)
+        k_vals = np.arange(-SS,SS-max(ll,nn)+0.5)
         ln_terms = ln_factors(k_vals) + ln_prefactor
         ln_term_max = ln_terms.max()
         terms = np.exp(ln_terms - ln_term_max)
         return ln_term_max + np.log(terms.sum())
 
     k_offset = SS % 1
-    k_vals_p = np.arange(1-k_offset, SS+1)
-    ln_terms_p = mm * np.log(k_vals_p) + ln_factors(+k_vals_p) + ln_prefactor
[…]

### 2018-09-15 soc_squeezing/cumulant_expansions.tex
changed convention for sum limits in expectation value w.r.t. X
indexed variables to clarify their dependencies
fixed small error and removed extra whitespace

@@
   \sp{S_+^p S_\z^q S_-^r, S_+^\ell S_\z^m S_-^n}_-
-  + \sum_\J \gamma_\J \D\p{\J;\O}
+  + \sum_\J \gamma_\J \D\p{\J; S_+^\ell S_-^m S_\z^n}
   \equiv \sum_{p,q,r} T^{\ell m n}_{pqr} S_+^p S_\z^q S_-^r,
@@
   $w\p{p,q,r}\ge W$ have a non-negligible contribution to operators of
-  interest.
-  \label{pt:limitation}
+  interest.\label{pt:limitation}
 \end{enumerate*}
@@
   &= \mu^m \f{N!}{2^N} \sum_{k=-S+\delta_{\mu,-1}\max\set{\ell,n}}
-  ^{S-\delta_{\mu,1}\max\set{\ell,n}} k^m
-  \f{\p{S-\mu k}!}{\p{S+\mu k}!\p{S-\mu k-\ell}!\p{S-\mu k-n}!} \\
-  &= \p{-1}^m \f{N!}{2^N} \sum_{k=-S+\max\set{\ell,n}}^S k^m
-  \f{\p{S+k}!}{\p{S-k}!\p{S+k-\ell}!\p{S+k-n}!}.
+  ^{S-\delta_{\mu,1}\max\set{\ell,n}}
+  \f{k^m \p{S-\mu k}!}
+  {\p{S+\mu k}!\p{S-\mu k-\ell}!\p{S-\mu k-n}!} \\
+  &= \f{N!}{2^N} \sum_{k=-S}^{S-\max\set{\ell,n}}
+  \f{k^m \p{S-k}!}{\p{S+k}!\p{S-k-\ell}!\p{S-k-n}!}.
 \end{align}
@@
   = \f{\p{N!}^2}{2^N\p{N-\ell}!\p{N-n}!}
-  \sp{\p{-S}^m \alpha + \f{\p{N-\ell}\p{N-n}}{N} \beta},
+  \sp{\p{-S}^m \alpha_{N\ell n}
+    + \f{\p{N-\ell}\p{N-n}}{N} \beta_{N\ell mn}},
 \end{align}
@@
 \begin{align}
-  \alpha
+  \alpha_{N\ell n}
   \equiv \F\p{\let\scriptstyle\textstyle
@@
 \begin{align}
-  \beta
+  \beta_{N\ell mn}
   \equiv \sum_{p=0}^{m-1} \p{-S}^p { m \choose p }

### 2018-09-15 soc_squeezing/cumulant_expansions.tex
fixed lots of (probably copy/pasted) typos with incorrect exponents on S_z

@@
 \end{align}
-The evolution of a general operator $S_+^\ell S_-^m S_\z^n$ under a
+The evolution of a general operator $S_+^\ell S_\z^m S_-^n$ under a
 Hamiltonian of the form in \eqref{eq:H_general} is given by
@@
   \sp{S_+^p S_\z^q S_-^r, S_+^\ell S_\z^m S_-^n}_-
-  + \sum_\J \gamma_\J \D\p{\J; S_+^\ell S_-^m S_\z^n}
+  + \sum_\J \gamma_\J \D\p{\J; S_+^\ell S_\z^m S_-^n}
   \equiv \sum_{p,q,r} T^{\ell m n}_{pqr} S_+^p S_\z^q S_-^r,
@@
   \sum_j \sigma_\z^{(j)} S_\mu^\ell
-  \p{\mu S_\z}^n S_\bmu^n \sigma_\z^{(j)}
+  \p{\mu S_\z}^m S_\bmu^n \sigma_\z^{(j)}
   &= 2\p{S-\ell-n} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n \notag \\
@@
   \sum_j \sigma_\z^{(j)} S_\mu^\ell
-  \p{\mu S_\z}^n S_\bmu^n \sigma_\mu^{(j)}
+  \p{\mu S_\z}^m S_\bmu^n \sigma_\mu^{(j)}
   &= \mu S_\mu^{\ell+1} \p{1+\mu S_\z}^m S_\bmu^n
@@
   \sum_j \sigma_\z^{(j)} S_\mu^\ell
-  \p{\mu S_\z}^n S_\bmu^n \sigma_\bmu^{(j)}
+  \p{\mu S_\z}^m S_\bmu^n \sigma_\bmu^{(j)}
   &= -\mu S_\mu^\ell \p{\mu S_\z}^m S_\bmu^{n+1}
-  + 2\mu \ell S_\mu^{\ell-1} \p{S+\mu S_\z}\p{-1+\mu S_\z}^m S_\bmu^n,
+  + 2\mu \ell S_\mu^{\ell-1} \p{S+\mu S_\z} \p{-1+\mu S_\z}^m S_\bmu^n,
   \label{eq:san_z_bmu} \allowdisplaybreaks \\
   \sum_j \sigma_\mu^{(j)} S_\mu^\ell
-  \p{\mu S_\z}^n S_\bmu^n \sigma_\mu^{(j)}
+  \p{\mu S_\z}^m S_\bmu^n \sigma_\mu^{(j)}
   &= n S_\mu^{\ell+1} \p{\mu S_\z}^m S_\bmu^{n-1}
@@
   \sum_j \sigma_\mu^{(j)} S_\mu^\ell
-  \p{\mu S_\z}^n S_\bmu^n \sigma_\bmu^{(j)}
+  \p{\mu S_\z}^m S_\bmu^n \sigma_\bmu^{(j)}
   &= S_\mu^\ell \p{S+\mu S_\z}\p{-1+\mu S_\z}^m S_\bmu^n,
@@
   \sum_j \sigma_\bmu^{(j)} S_\mu^\ell
-  \p{\mu S_\z}^n S_\bmu^n \sigma_\mu^{(j)}
+  \p{\mu S_\z}^m S_\bmu^n \sigma_\mu^{(j)}
   &= S_\mu^\ell \p{S - \ell - n - \mu S_\z}
@@
 \begin{align}
-  \sp{S_\z, S_\mu^\ell \p{\mu S_\z}^n S_\bmu^n}_+
+  \sp{S_\z, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
   = S_\z S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n
@@
 \begin{align}
-  \D\p{\sigma_\bmu; S_\mu^\ell \p{\mu S_\z}^n S_\bmu^n}
+  \D\p{\sigma_\bmu; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
   = S_\mu^\ell \p{S+\mu S_\z}\p{-1+\mu S_\z}^m S_\bmu^n
@@
 \begin{align}
-  \D\p{\sigma_\mu; S_\mu^\ell \p{\mu S_\z}^n S_\bmu^n}
+  \D\p{\sigma_\mu; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
   &= S_\mu^\ell \p{S - \ell - n - \mu S_\z} \p{1+\mu S_\z}^m S_\bmu^n
@@
 \begin{align}
-  \D\p{\sigma_\z/\sqrt2; S_\mu^\ell \p{\mu S_\z}^n S_\bmu^n}
+  \D\p{\sigma_\z/\sqrt2; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
   = -\p{\ell+n} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n
@@
 \begin{align}
-  \sp{S_\mu, S_\mu^\ell \p{\mu S_\z}^n S_\bmu^n}_+
+  \sp{S_\mu, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
   = S_\mu^{\ell+1} \sp{S_\z^m+\p{1+\mu S_\z}^m} S_\bmu^n
@@
   \sum_j \sigma_\z^{(j)} S_\mu^\ell
-  \p{\mu S_\z}^n S_\bmu^n \sigma_\mu^{(j)}
+  \p{\mu S_\z}^m S_\bmu^n \sigma_\mu^{(j)}
   - \f{\mu}{2} \sp{S_\mu, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
@@
   \sum_j \sigma_\mu^{(j)} S_\mu^\ell
-  \p{\mu S_\z}^n S_\bmu^n \sigma_\z^{(j)}
-  + \f{\mu}{2} \sp{S_\mu, S_\mu^\ell \p{\mu S_\z}^n S_\bmu^n}_+
+  \p{\mu S_\z}^m S_\bmu^n \sigma_\z^{(j)}
+  + \f{\mu}{2} \sp{S_\mu, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
   = K_{\ell mn;\mu} + M_{\ell mn;\mu}
@@
 \begin{align}
-  \D\p{\gamma; S_\mu^\ell \p{\mu S_\z}^n S_\bmu^n}
+  \D\p{\gamma; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
   = \sum_{X\in\set{\z,+,-}} \abs{\gamma_X}^2
@@
 Note that the sum $\ell+m+n$ for operators
-$S_\mu^\ell \p{\mu S_\z}^n S_\bmu^n$ does not grow under this type of
+$S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$ does not grow under this type of
 decoherence.
@@
 Note that the sum $\ell+m+n$ for operators
-$S_\mu^\ell \p{\mu S_\z}^n S_\bmu^n$ grows by one if $\Gamma_\mu\ne0$
+$S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$ grows by one if $\Gamma_\mu\ne0$
 or $\Gamma_\bmu\ne0$, and does not grow otherwise.

### 2018-09-26 soc_squeezing/cumulant_expansions.tex
removed incorrect interpretation of a Stirling number
minor changes to text

@@
 track of in order to compute the expectation value
-$\bk{S_+^\ell S_\z^m S_-^n}$.  Furthermore, after computing the
-expectation values of all relevant operators at $t=0$, time evolution
-of $\bk{S_+^\ell S_\z^m S_-^n}$ takes only $\O\p{M}$ CPU time.
-Finally, the relation
+$\bk{S_+^\ell S_\z^m S_-^n}$.  Furthermore, after computing all
+relevant matrix elements $T^{\ell mn}_{pqr;k}$ and expectation values
+$\bk{S_+^p S_\z^q S_-^r}_{t=0}$, computing
+$\bk{S_+^\ell S_\z^m S_-^n}$ takes only a short time independent of
+system size.  Finally, the relation
 $\p{S_+^\ell S_\z^m S_-^n}^\dag=S_+^n S_\z^m S_-^\ell$ can be
-exploited to cut both the number of expectation values
-$\bk{S_+^p S_\z^r S_-^q}_{t=0}$ and the number of matrix elements
-$T^{\ell mn}_{pqr}$ which we need to compute roughly in half.
+exploited to cut both the number of matrix elements
+$T^{\ell mn}_{pqr}$ and expectation values
+$\bk{S_+^p S_\z^r S_-^q}_{t=0}$ that we need to compute roughly in
+half.
 
@@
   &&
-  \Omega\p{t} = \beta\omega\cos\p{\omega t} S_\x,
-\end{align}
-where $\beta$ is the modulation index of the driving field, and the
-drive frequency $\omega\gg N\chi$.  Moving into the rotating frame of
-$\Omega\p{t}S_\x$ subtracts this term from the Hamiltonian, and
-transforms operators $\O$ as
+  \Omega\p{t} = \beta\omega \cos\p{\omega t},
+\end{align}
+where $\beta$ is the modulation index of the drive, and the drive
+frequency $\omega\gg N\chi$, with $N$ the total number of spins.
+Moving into the rotating frame of $\Omega\p{t}S_\x$ subtracts this
+term from the Hamiltonian, and transforms operators $\O$ as
 \begin{align}
@@
   &&
-  U\p{t} \equiv \exp\p{-i\int_0^t\d\tau~\Omega\p{\tau}S_\x}
+  U\p{t} \equiv \exp\sp{-i\int_0^t\d\tau~\Omega\p{\tau}S_\x}
   = \exp\sp{-i\beta\sin\p{\omega t}S_\x},
@@
 In particular, the operators $\tilde S_\pm\equiv-S_\z\pm iS_\y$
-transform as
+transform simply as
 \begin{align}
@@
 \end{align}
-For large drive frequencies $\omega$, we can generally make a secular
-approximation to say
-\begin{align}
-  e^{\pm i\beta\sin\p{\omega t}}
-  = \sum_{n=-\infty}^\infty J_n\p{\pm\beta} e^{in\omega t}
-  \approx J_0\p{\pm\beta} = J_0\p{\beta},
+For any operator $\O$ and drive frequency $\omega\gg\norm{\O}$, where
+$\norm{\O}$ denotes the operator norm of $\O$, we can generally make a
+secular approximation to say
+\begin{align}
+  e^{\pm im\beta\sin\p{\omega t}} \O
+  = \sum_{n=-\infty}^\infty J_n\p{\pm m\beta} e^{in\omega t} \O
+  \approx J_0\p{\pm m\beta} \O
+  = J_0\p{m\beta} \O,
 \end{align}
 where $J_n$ is the $n$-th order Bessel function of the first kind.
-Expanding $S_\z^2=\f14\p{\tilde S_++\tilde S_-}^2$, one can thus work
-out that the effective Hamiltonian in the rotating frame of the drive
-is
+Expanding $S_\z^2=\f14\p{\tilde S_+ + \tilde S_-}^2$, one can thus
+work out that the effective Hamiltonian in the rotating frame of the
+drive is
 \begin{align}
   H_{\t{eff}}
-  \approx \f{\chi}{2} \p{\sp{J_0\p{2\beta}+1} S_\z^2
-  - \sp{J_0\p{2\beta}-1} S_\y^2},
-\end{align}
-where choosing $\beta$ such that $J_0\p{2\beta}=\pm1/3$ will give us
-$H_{\t{TAT}}^{\z,\y}$ or $H_{\t{TAT}}^{\y,\x}$ up to irrelevant terms.
-In a similar spirit, one can work out that
+  \approx \f{\chi}{2} \p{\sp{1+J_0\p{2\beta}} S_\z^2
+    + \sp{1-J_0\p{2\beta}} S_\y^2}.
+\end{align}
+Driving with a modulation index $\beta$ for which
+$J_0\p{2\beta}=\pm1/3$ then gives us the effective two-axis twisting
+Hamiltonians
+\begin{align}
+  H_{\t{eff}}^{(+)}
+  = \f{\chi}{3} \p{2 S_\z^2 + S_\y^2}
+  \simeq \f{\chi}{3} \p{S_\z^2 - S_\x^2},
+  &&
+  H_{\t{eff}}^{(-)}
+  = \f{\chi}{3} \p{S_\z^2 + 2 S_\y^2}
+  \simeq \f{\chi}{3} \p{S_y^2 - S_\x^2},
+\end{align}
+where $\simeq$ denotes equality up to overall factors of
+$\v S^2=S_\z^2+S_\x^2+S_\y^2$ which are irrelevant in the absence of
+coherent coupling between states with different net spin.  In a
+similar spirit, one can work out that single-spin operators transverse
+to the $\x$-axis transform as
 \begin{align}
@@
 where ${ n \brack k }$ is an unsigned Stirling number of the first
-kind, i.e.~the number of ways to pick a set of $k$ elements from a set
-of $n$ elements, and finally
+kind, and finally
 \begin{align}

### 2018-09-28 soc_squeezing/cumulant_expansions.tex
removed unnecessary brackets
fixed a few minor errors and clarified a bit of text

@@
   r \equiv \gamma_+ \gamma_-,
-  &
+  &&
   \Gamma \equiv \gamma_\z + \lambda,
@@
 operators $S_+^\ell S_\z^m S_-^n$ to spin operators
-$S_+^p S_\z^q S_-^r$ with higher weight, i.e.~with $p+q+r>\ell+m+n$.
-In practice, keeping track of this operator growth will eventually
-require more computational resources than are available, so we must
-truncate it at some point by taking
+$S_+^p S_\z^q S_-^r$ with higher ``weight'', i.e.~with
+$p+q+r>\ell+m+n$, which signifies the growth of many-body
+correlations.  In practice, keeping track of this operator growth will
+eventually require more computational resources than are available, so
+we must truncate it at some point by taking
 \begin{align}
@@
   of operators can be very time-consuming,
-\item the weight function $w$ must be chosen carefully, as the optimal
-  weight function is generally system-dependent,
+\item the weight function $w$ may need to be chosen carefully, as the
+  optimal weight function is generally system-dependent,
 \item simulation results can only be trusted up to the time at which
@@
   f_{np\ell;\mu}\p{k_{p-2},1} \\
-  &= \f1{r!} \sum_{\p{\v k,p-r,\ell}}
+  &= \f1{\p{r-1}!} \sum_{\p{\v k,p-r,\ell}}
   \prod_{j=r}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}}
@@
   \tilde Z_{qr\ell m;\mu}^{(k)}
-  &\equiv \p{\sp{\ell-k}+\mu S_\z}^q
-  Z_{r\ell;\mu}^{(k)} \p{\sp{r-k}+\mu S_\z}^m \\
+  &\equiv \p{\ell-k+\mu S_\z}^q
+  Z_{r\ell;\mu}^{(k)} \p{r-k+\mu S_\z}^m \\
   &= \sum_{a=0}^k \zeta_{r\ell}^{ka}

### 2018-09-29 soc_squeezing/cumulant_expansions.tex
added text about computing the product of two k-body spin operators

@@
 
+\section{Changing operator bases}
+
+We wish to change our basis for all exchange-symmetric operators on a
+system of $N$ qubits from $S_+^\ell S_\z^m S_-^n$ to
+\begin{align}
+  \S_{\v m} \equiv \sum_{\p{\v j}} P_{\v m\v j}
+\end{align}
+where
+\begin{align}
+  \v m \equiv \p{m_+,m_\z,m_-},
+  &&
+  \v j \equiv \p{j_1^+,j_2^+,\cdots j_{m_+}^+,
+    j_1^\z,j_2^\z,\cdots j_{m_\z}^\z,
+    j_1^-,j_2^-,\cdots j_{m_-}^-},
+\end{align}
+and for shorthand we define
+\begin{align}
+  \sum_{\p{\v j}} X_{\v j} \equiv
+  \sum_{j_a^\mu=j_b^\nu \iff \p{a,\mu}=\p{b,\nu}} X_{\v j},
+  &&
+  P_{\v m\v j} \equiv
+  \prod_\mu \prod_{a=1}^{m_\mu}
+  \sigma_\mu^{\p{j_a^\mu}}.
+\end{align}
+In words, $\S_{\v m}$ is a $\p{m_++m_\z+m_-}$-spin operator with
+$m_\mu$ single-spin operators of type $\sigma_\mu$ addressing the
+spins indexed by $j_1^\mu,j_2^\mu,\cdots j_{m_\mu}^\mu$.  We now
+expand a product of basis operators $\S_{\v m}$ and $\S_{\v n}$ into
+terms that have exactly $s$ spins addressed by both operators:
+\begin{align}
+  \S_{\v m} \S_{\v n}
+  = \sum_{s\ge0} \sum_{\p{\v j,\v k,s}} P_{\v m\v j} P_{\v n\v k},
+  &&
+  \sum_{\p{\v j,\v k,s}} X_{\v j\v k} \equiv
+  \sum_{\substack{\p{\v j},\p{\v k} \\
+      \abs{\set{j_\alpha^\mu}\cap\set{k_\beta^\nu}}=s}}
+  X_{\v j\v k},
+  \label{eq:SS_PP}
+\end{align}
+where the sum over $s$ terminates after $s = \sum_\mu\p{m_\mu+n_\mu}$.
+Collecting terms in which $r_{\mu\nu}$ of the $\sigma_\mu$ operators
+in $P_{\v m\v j}$ address the same spin as a $\sigma_\nu$ operator in
+$P_{\v n\v k}$, we have
+\begin{align}
+  \S_{\v m}\S_{\v n}
+  = \sum_{s\ge0} \sum_{\p{\v m,\v n,\v r,\v J,\v K}}
+  f_{\v m\v n\v r} Q_{\v r\v J} \tilde Q_{\v r\v K},
+  \label{eq:SS_PQ}
+\end{align}
+where $\set{J_a^\mu}$ and $\set{K_a^{\mu\nu}}$ are respectively the
+indices of spins addressed by one and two single-spin operators in
+each term of \eqref{eq:SS_PP}; the sum over
+$\p{\v m,\v n,\v r,\v J,\v K}$ implies a sum over all possible values
+of $\v r,\v J,\v K$ with the restrictions
+\begin{align}
+  \sum_{\mu,\nu} r_{\mu\nu} = s,
+  &&
+  0 \le \sum_\nu r_{\mu\nu} \le m_\mu,
+  &&
+  0 \le \sum_\mu r_{\mu\nu} \le n_\nu,
+\end{align}
+\begin{align}
+  \abs{\set{J_a^\mu}\cup\set{K_a^{\mu\nu}}}
+  = \sum_\mu\p{m_\mu+n_\mu} - s;
+\end{align}
+the factor $f_{\v m\v n\v r}$ accounts for the number of ways to pair
+operators in $P_{\v m\v j}$ and $p_{\v n\v k}$ as prescribed by
+$\v r$, or
+\begin{align}
+  f_{\v m\v n\v r} \equiv \sp{\prod_{\mu,\nu}
+    { m_\mu - \sum_{\rho>\nu} r_{\mu\rho} \choose r_{\mu\nu} }
+    { n_\mu - \sum_{\rho>\nu} r_{\rho\mu} \choose r_{\nu\mu} }}
+  \sp{\prod_{\mu\nu} r_{\mu\nu}!},
+\end{align}
+and
+\begin{align}
+  Q_{\v r\v J} \equiv \prod_\mu
+  \prod_{a=1}^{\abs{\set{J_b^\mu}}} \sigma_\mu^{(J_a^\mu)},
+  &&
+  \tilde Q_{\v r\v K} \equiv \prod_{\mu,\nu}
+  \prod_{a=1}^{\abs{\set{K_b^{\mu\nu}}}}
+  \sigma_\mu^{(K_a^{\mu\nu})} \sigma_\nu^{(K_a^{\mu\nu})}.
+\end{align}
+
 \end{document}

### 2018-09-29 soc_squeezing/cumulant_expansions.tex
simplified some expressions in the appendix on changing operator bases

@@
   &&
-  P_{\v m\v j} \equiv
-  \prod_\mu \prod_{a=1}^{m_\mu}
-  \sigma_\mu^{\p{j_a^\mu}}.
+  P_{\v m\v j} \equiv \prod_{\mu,a} \sigma_\mu^{\p{j_a^\mu}}.
 \end{align}
@@
 \begin{align}
+  r_{\mu\nu} \ge 0 ~\forall~ \mu,\nu,
+  &&
   \sum_{\mu,\nu} r_{\mu\nu} = s,
   &&
-  0 \le \sum_\nu r_{\mu\nu} \le m_\mu,
-  &&
-  0 \le \sum_\mu r_{\mu\nu} \le n_\nu,
-\end{align}
-\begin{align}
-  \abs{\set{J_a^\mu}\cup\set{K_a^{\mu\nu}}}
-  = \sum_\mu\p{m_\mu+n_\mu} - s;
+  \sum_\nu r_{\mu\nu} \le m_\mu,
+  &&
+  \sum_\mu r_{\mu\nu} \le n_\nu,
+\end{align}
+\begin{align}
+  \abs{\set{J_a^\mu}} = \sum_\mu\p{m_\mu+n_\mu} - 2 s,
+  &&
+  \abs{\set{K_a^{\mu\nu}}} = s,
+  &&
+  \set{J_a^\mu} \cap \set{K_a^{\mu\nu}} = \set{};
 \end{align}
 the factor $f_{\v m\v n\v r}$ accounts for the number of ways to pair
-operators in $P_{\v m\v j}$ and $p_{\v n\v k}$ as prescribed by
+operators in $P_{\v m\v j}$ and $P_{\v n\v k}$ as prescribed by
 $\v r$, or
 \begin{align}
-  f_{\v m\v n\v r} \equiv \sp{\prod_{\mu,\nu}
+  f_{\v m\v n\v r}
+  &\equiv \sp{\prod_{\mu,\nu}
     { m_\mu - \sum_{\rho>\nu} r_{\mu\rho} \choose r_{\mu\nu} }
     { n_\mu - \sum_{\rho>\nu} r_{\rho\mu} \choose r_{\nu\mu} }}
-  \sp{\prod_{\mu\nu} r_{\mu\nu}!},
-\end{align}
-and
-\begin{align}
-  Q_{\v r\v J} \equiv \prod_\mu
-  \prod_{a=1}^{\abs{\set{J_b^\mu}}} \sigma_\mu^{(J_a^\mu)},
-  &&
-  \tilde Q_{\v r\v K} \equiv \prod_{\mu,\nu}
-  \prod_{a=1}^{\abs{\set{K_b^{\mu\nu}}}}
+  \sp{\prod_{\mu,\nu} r_{\mu\nu}!} \\
+  &= \sp{\prod_\mu \p{m_\mu}_{\sum_\rho r_{\mu\rho}}
+    \p{n_\mu}_{\sum_\rho r_{\rho\mu}}}
+  \sp{\prod_{\mu\nu} r_{\mu\nu}!}^{-1},
+\end{align}
+where $\p{n}_k\equiv\prod_{j=0}^{k-1}\p{n-j}$ is a falling factorial;
+and finally
+\begin{align}
+  Q_{\v r\v J} \equiv \prod_{\mu,a} \sigma_\mu^{(J_a^\mu)},
+  &&
+  \tilde Q_{\v r\v K} \equiv \prod_{\mu,\nu,a}
   \sigma_\mu^{(K_a^{\mu\nu})} \sigma_\nu^{(K_a^{\mu\nu})}.

### 2018-09-29 soc_squeezing/cumulant_expansions.tex
improved some notation

--- 2018-09-28 soc_squeezing/cumulant_expansions.tex
@@
 %%% symbols, notations, etc.
-\usepackage{physics,braket,bm,commath,amssymb} % physics and math
+\usepackage{physics,braket,bm,commath,amssymb}
 \renewcommand{\t}{\text} % text in math mode
@@
 
+\section{Changing operator bases}
+
+We wish to change our basis for all exchange-symmetric operators on a
+system of $N$ qubits from $S_+^\ell S_\z^m S_-^n$ to
+\begin{align}
+  \S_{\v m} \equiv \sum_{\p{\v j;\v m}} P_{\v j},
+  &&
+  P_{\v j} \equiv \prod_{\mu,a} \sigma_\mu^{\p{j_a^\mu}},
+\end{align}
+where $\v m \equiv \p{m_+,m_\z,m_-}$; $\v j$ is an ordered list of
+indices for the spins addressed by $P_{\v j}$; and the sum over
+$\p{\v j;\v m}$ denotes a sum over all possible $\v j$ for which all
+indices in $\v j$ are distinct and $\abs{\set{j_a^\mu}} = m_\mu$.  In
+words, $\S_{\v m}$ is an exchange-symmetric $\p{m_++m_\z+m_-}$-spin
+operator with $m_\mu$ single-spin operators of type $\sigma_\mu$, and
+$P_{\v j}$ is a single term in $\S_{\v m}$ which addresses the spins
+indexed by $\v j$.
+
+We now expand a product of basis operators $\S_{\v m}$ and $\S_{\v n}$
+into terms that have exactly $s$ spins addressed by both operators:
+\begin{align}
+  \S_{\v m} \S_{\v n}
+  = \sum_{s\ge0} \sum_{\p{\v j,\v k;\v m,\v n,s}} P_{\v j} P_{\v k},
+  &&
+  \sum_{\p{\v j,\v k;\v m,\v n,s}} X \equiv
+  \sum_{\substack{\p{\v j;\v m},\p{\v k;\v n} \\
+      \abs{\set{j_\alpha^\mu}\cap\set{k_\beta^\nu}}=s}} X,
+  \label{eq:SS_PP}
+\end{align}
+where the sum over $s$ terminates after $s = \sum_\mu\p{m_\mu+n_\mu}$.
+Collecting terms in which $r_{\mu\nu}$ of the $\sigma_\mu$ operators
+in $P_{\v j}$ address the same spin as a $\sigma_\nu$ operator in
+$P_{\v k}$, we have
+\begin{align}
+  \S_{\v m}\S_{\v n}
+  = \sum_{s\ge0} \sum_{\p{\v r,\v J,\v K;\v m,\v n,s}}
+  f_{\v m\v n\v r} Q_{\v J} \tilde Q_{\v K},
+  \label{eq:SS_PQ}
+\end{align}
+where $\v J$ and $\v K$ are respectively ordered lists of indices for
+spins addressed by one and two single-spin operators in each term of
+the expansion for $\S_{\v m} \S_{\v n}$ in \eqref{eq:SS_PP}; the sum
+over $\p{\v r,\v J,\v K;\v m,\v n,s}$ implies a sum over all possible
+values of $\v r,\v J,\v K$ with the restrictions
+\begin{align}
+  r_{\mu\nu} \ge 0 ~\forall~ \mu,\nu,
+  &&
+  \sum_\nu r_{\mu\nu} \le m_\mu,
+  &&
+  \sum_\mu r_{\mu\nu} \le n_\nu,
+  &&
+  \sum_{\mu,\nu} r_{\mu\nu} = s,
+\end{align}
+\begin{align}
+  \abs{\set{J_a^\mu}} = \sum_\mu\p{m_\mu+n_\mu} - 2 s,
+  &&
+  \abs{\set{K_a^{\mu\nu}}} = s,
+  &&
+  \set{J_a^\mu} \cap \set{K_a^{\mu\nu}} = \set{};
+\end{align}
+the factor $f_{\v m\v n\v r}$ counts the number of ways to pair
+operators in $P_{\v j}$ and $P_{\v k}$ as prescribed by $\v r$, or
+\begin{align}
+  f_{\v m\v n\v r}
+  &\equiv \sp{\prod_{\mu,\nu}
+    { m_\mu - \sum_{\rho>\nu} r_{\mu\rho} \choose r_{\mu\nu} }
+    { n_\mu - \sum_{\rho>\nu} r_{\rho\mu} \choose r_{\nu\mu} }}
+  \sp{\prod_{\mu,\nu} r_{\mu\nu}!} \\
+  &= \sp{\prod_\mu \p{m_\mu}_{\sum_\rho r_{\mu\rho}}
+    \p{n_\mu}_{\sum_\rho r_{\rho\mu}}}
+  \sp{\prod_{\mu\nu} r_{\mu\nu}!}^{-1},
+\end{align}
+where $\p{n}_k\equiv\prod_{j=0}^{k-1}\p{n-j}$ is a falling factorial;
+and finally
+\begin{align}
+  Q_{\v J} \equiv \prod_{\mu,a} \sigma_\mu^{(J_a^\mu)},
+  &&
+  \tilde Q_{\v K} \equiv \prod_{\mu,\nu,a}
+  \sigma_\mu^{(K_a^{\mu\nu})} \sigma_\nu^{(K_a^{\mu\nu})}.
+\end{align}
+
 \end{document}

### 2018-09-29 soc_squeezing/cumulant_expansions.tex
fixed typo
clarified restrictions on sum over r,J,K
simplified a general product of two operators

@@
   = \sum_{s\ge0} \sum_{\p{\v r,\v J,\v K;\v m,\v n,s}}
-  f_{\v m\v n\v r} Q_{\v J} \tilde Q_{\v K},
+  g_{\v m\v n\v r} Q_{\v J} \tilde Q_{\v K},
   \label{eq:SS_PQ}
@@
 \begin{align}
-  \abs{\set{J_a^\mu}} = \sum_\mu\p{m_\mu+n_\mu} - 2 s,
-  &&
-  \abs{\set{K_a^{\mu\nu}}} = s,
+  \abs{\set{J_a^\mu}}
+  = m_\mu + n_\mu - \sum_\rho\p{r_{\mu\rho}+r_{\rho\mu}},
+  &&
+  \abs{\set{K_a^{\mu\nu}}} = r_{\mu\nu},
   &&
@@
 \end{align}
-the factor $f_{\v m\v n\v r}$ counts the number of ways to pair
+the factor $g_{\v m\v n\v r}$ counts the number of ways to pair
 operators in $P_{\v j}$ and $P_{\v k}$ as prescribed by $\v r$, or
 \begin{align}
-  f_{\v m\v n\v r}
+  g_{\v m\v n\v r}
   &\equiv \sp{\prod_{\mu,\nu}
-    { m_\mu - \sum_{\rho>\nu} r_{\mu\rho} \choose r_{\mu\nu} }
-    { n_\mu - \sum_{\rho>\nu} r_{\rho\mu} \choose r_{\nu\mu} }}
+    { m_\mu - \sum_{\rho<\nu} r_{\mu\rho} \choose r_{\mu\nu} }
+    { n_\mu - \sum_{\rho<\nu} r_{\rho\mu} \choose r_{\nu\mu} }}
   \sp{\prod_{\mu,\nu} r_{\mu\nu}!} \\
@@
     \p{n_\mu}_{\sum_\rho r_{\rho\mu}}}
-  \sp{\prod_{\mu\nu} r_{\mu\nu}!}^{-1},
+  \sp{\prod_{\mu,\nu} r_{\mu\nu}!}^{-1},
 \end{align}
@@
   \tilde Q_{\v K} \equiv \prod_{\mu,\nu,a}
-  \sigma_\mu^{(K_a^{\mu\nu})} \sigma_\nu^{(K_a^{\mu\nu})}.
-\end{align}
+  \sigma_\mu^{(K_a^{\mu\nu})} \sigma_\nu^{(K_a^{\mu\nu})}
+  = \prod_{\mu,\nu,a} \sum_\rho
+  \eta_{\mu\nu\rho} \sigma_\rho^{(K_a^{\mu\nu})},
+  \label{eq:Q_JK}
+\end{align}
+where $\eta_{\mu\nu\rho}$ is defined by the product
+$\sigma_\mu\sigma_\nu\equiv\sum_\rho\eta_{\mu\nu\rho}\sigma_\rho$.
+Note that unlike the sums over $\mu$ and $\nu$ in the work above, the
+sum over $\rho$ in \eqref{eq:Q_JK} now includes an index for the
+identity operator $\sigma_0\equiv\1$.
 

### 2018-10-01 soc_squeezing/cumulant_expansions.tex
removed redundant notation
fixed a few errors and futher simplified the product of two k-body operators

--- 2018-09-29 soc_squeezing/cumulant_expansions.tex
@@
   = \sum_{s\ge0} \sum_{\p{\v r,\v J,\v K;\v m,\v n,s}}
-  f_{\v m\v n\v r} Q_{\v J} \tilde Q_{\v K},
+  f_{\v m\v n\v r} P_{\v J} Q_{\v K},
   \label{eq:SS_PQ}
@@
 the expansion for $\S_{\v m} \S_{\v n}$ in \eqref{eq:SS_PP}; the sum
-over $\p{\v r,\v J,\v K;\v m,\v n,s}$ implies a sum over all possible
-values of $\v r,\v J,\v K$ with the restrictions
+over $\p{\v r,\v J,\v K;\v m,\v n,s}$ denotes a sum over all values of
+$\v r,\v J,\v K$ with the restrictions
 \begin{align}
@@
 \begin{align}
-  \abs{\set{J_a^\mu}} = \sum_\mu\p{m_\mu+n_\mu} - 2 s,
-  &&
-  \abs{\set{K_a^{\mu\nu}}} = s,
-  &&
-  \set{J_a^\mu} \cap \set{K_a^{\mu\nu}} = \set{};
+  \abs{\set{J_a^\mu}}
+  = m_\mu + n_\mu - \sum_\rho\p{r_{\mu\rho}+r_{\rho\mu}},
+  &&
+  \abs{\set{K_a^{\mu\nu}}} = r_{\mu\nu},
+  &&
+  \set{J_a^\lambda} \cap \set{K_b^{\mu\nu}} = \set{};
 \end{align}
@@
   &\equiv \sp{\prod_{\mu,\nu}
-    { m_\mu - \sum_{\rho>\nu} r_{\mu\rho} \choose r_{\mu\nu} }
-    { n_\mu - \sum_{\rho>\nu} r_{\rho\mu} \choose r_{\nu\mu} }}
+    { m_\mu - \sum_{\rho<\nu} r_{\mu\rho} \choose r_{\mu\nu} }
+    { n_\mu - \sum_{\rho<\nu} r_{\rho\mu} \choose r_{\nu\mu} }}
   \sp{\prod_{\mu,\nu} r_{\mu\nu}!} \\
@@
     \p{n_\mu}_{\sum_\rho r_{\rho\mu}}}
-  \sp{\prod_{\mu\nu} r_{\mu\nu}!}^{-1},
+  \sp{\prod_{\mu,\nu} r_{\mu\nu}!}^{-1},
 \end{align}
 where $\p{n}_k\equiv\prod_{j=0}^{k-1}\p{n-j}$ is a falling factorial;
-and finally
-\begin{align}
-  Q_{\v J} \equiv \prod_{\mu,a} \sigma_\mu^{(J_a^\mu)},
-  &&
-  \tilde Q_{\v K} \equiv \prod_{\mu,\nu,a}
+and finally, similarly to the definition of $P_{\v j}$ in
+\eqref{eq:SS_PP} we have
+\begin{align}
+  Q_{\v K} \equiv \prod_{\mu,\nu,a}
   \sigma_\mu^{(K_a^{\mu\nu})} \sigma_\nu^{(K_a^{\mu\nu})}.
 \end{align}
+Expanding the product
+$\sigma_\mu\sigma_\nu\equiv\sum_\rho\eta_{\mu\nu\rho}\sigma_\rho$,
+where unlike the sums over $\mu,\nu$ in most of the work above the sum
+over $\rho$ here includes an index for the identity operator
+$\sigma_0\equiv\1$, we can simplify
+\begin{align}
+  Q_{\v K}
+  = \prod_{\mu,\nu,a} \sum_\rho
+  \eta_{\mu\nu\rho} \sigma_\rho^{(K_a^{\mu\nu})},
+\end{align}
+and collect terms for each $\p{\mu,\nu}$ that are unique up to a
+permutation of indices and have $c_\rho$ single-spin operators of
+the type $\sigma_\rho$, giving us
+\begin{align}
+  Q_{\v K}
+  \cong \prod_{\mu,\nu} \sum_{\sum_\lambda c_\lambda=r_{\mu\nu}}
+  g^{\mu\nu}_{\v c\v r} \prod_\rho \p{\eta_{\mu\nu\rho}}^{c_\rho}
+  \prod_{a=1}^{c_\rho} \sigma_\rho^{(K^{\mu\nu\rho}_{\v ca})},
+\end{align}
+where $\cong$ denotes equality under the sum in \eqref{eq:SS_PQ} (as
+we have freely permuted spin indices and introduced explicit
+dependence on $\v r$ for convenience); the factor
+$g^{\mu\nu}_{\v c\v r}$ counts the number of ways to choose operators
+in $Q_{\v K}$ according to $\v c$ (i.e.~the number of choices that are
+equivalent up to permutation of indices in $\v K$), or
+\begin{align}
+  g^{\mu\nu}_{\v c\v r} \equiv \prod_\rho
+  { r_{\mu\nu} - \sum_{\lambda<\rho} c_\lambda \choose c_\rho }
+  = \f{r_{\mu\nu}!}{\prod_\rho c_\rho!};
+\end{align}
+and
+\begin{align}
+  K^{\mu\nu\rho}_{\v ca}
+  \equiv K^{\mu\nu}_{\sum_{\lambda<\rho}c_\lambda+a}
+\end{align}
+indexes the $a$-th operator of type $\sigma_\rho$, given a choice of
+$\v c$.
 

### 2018-10-01 soc_squeezing/cumulant_expansions.tex
fixed removed incorrect and untested analytical work, and wrote new, tested analytical work on product of operators

@@
   &&
-  \abs{\set{K_a^{\mu\nu}}} = r_{\mu\nu},
+  \abs{\set{K^{\mu\nu}_a}} = r_{\mu\nu},
   &&
@@
   Q_{\v K} \equiv \prod_{\mu,\nu,a}
-  \sigma_\mu^{(K_a^{\mu\nu})} \sigma_\nu^{(K_a^{\mu\nu})}.
+  \sigma_\mu^{(K^{\mu\nu}_a)} \sigma_\nu^{(K^{\mu\nu}_a)}.
 \end{align}
 Expanding the product
-$\sigma_\mu\sigma_\nu\equiv\sum_\rho\eta_{\mu\nu\rho}\sigma_\rho$,
-where unlike the sums over $\mu,\nu$ in most of the work above the sum
-over $\rho$ here includes an index for the identity operator
-$\sigma_0\equiv\1$, we can simplify
+$\sigma_\mu\sigma_\nu=\sum_\rho\eta_{\mu\nu\rho}\sigma_\rho$, where
+unlike the sums over $\mu,\nu$ in most of the work above, the sum over
+$\rho$ here includes an index for the identity operator
+$\sigma_0\equiv\1$, we can expand and collect terms to get
 \begin{align}
@@
   = \prod_{\mu,\nu,a} \sum_\rho
-  \eta_{\mu\nu\rho} \sigma_\rho^{(K_a^{\mu\nu})},
-\end{align}
-and collect terms for each $\p{\mu,\nu}$ that are unique up to a
-permutation of indices and have $c_\rho$ single-spin operators of
-the type $\sigma_\rho$, giving us
+  \eta_{\mu\nu\rho} \sigma_\rho^{(K^{\mu\nu}_a)}
+  = \sum_{\v\rho} \prod_{\mu,\nu,a} \eta_{\mu\nu\rho^{\mu\nu}_a}
+  \sigma_{\rho^{\mu\nu}_a}^{(K^{\mu\nu}_a)},
+\end{align}
+where we have distributed a product of sums into a sum of products.
+Collecting together operators of type $\sigma_\kappa$, we then have
 \begin{align}
   Q_{\v K}
-  \cong \prod_{\mu,\nu} \sum_{\sum_\lambda c_\lambda=r_{\mu\nu}}
-  g^{\mu\nu}_{\v c\v r} \prod_\rho \p{\eta_{\mu\nu\rho}}^{c_\rho}
-  \prod_{a=1}^{c_\rho} \sigma_\rho^{(K^{\mu\nu\rho}_{\v ca})},
-\end{align}
-where $\cong$ denotes equality under the sum in \eqref{eq:SS_PQ} (as
-we have freely permuted spin indices and introduced explicit
-dependence on $\v r$ for convenience); the factor
-$g^{\mu\nu}_{\v c\v r}$ counts the number of ways to choose operators
-in $Q_{\v K}$ according to $\v c$ (i.e.~the number of choices that are
-equivalent up to permutation of indices in $\v K$), or
-\begin{align}
-  g^{\mu\nu}_{\v c\v r} \equiv \prod_\rho
-  { r_{\mu\nu} - \sum_{\lambda<\rho} c_\lambda \choose c_\rho }
-  = \f{r_{\mu\nu}!}{\prod_\rho c_\rho!};
-\end{align}
-and
-\begin{align}
-  K^{\mu\nu\rho}_{\v ca}
-  \equiv K^{\mu\nu}_{\sum_{\lambda<\rho}c_\lambda+a}
-\end{align}
-indexes the $a$-th operator of type $\sigma_\rho$, given a choice of
-$\v c$.
+  = \sum_{\v\rho} \sp{\prod_{\mu,\nu,\kappa}
+    \p{\eta_{\mu\nu\kappa}}^{\tilde\rho_\kappa}}
+  \prod_{\kappa,a} \sigma_{\kappa}^{(\tilde K^{\v\rho\kappa}_a)},
+\end{align}
+where $\tilde\rho_\kappa$ is the number of elements in $\v\rho$ equal
+to $\kappa$, and $\tilde K^{\v\rho\kappa}_a$ indexes the $a$-th
+operator of type $\sigma_\kappa$ for each choice of $\v\rho$, such
+that for any fixed $\v\rho$ we have
+$\set{\tilde K^{\v\rho\kappa}_a}=\set{K^{\mu\nu}_b}$.
 

### 2018-10-01 soc_squeezing/scripts/test_product.py

@@ -41,6 +41,14 @@ def spin_op(mu):
     if mu == 2: return sm
     if mu == 3: return I2
 
+def PP(jj,mm):
+    jj_vals = partition(jj,mm)
+    op_list = [ I2 ] * N
+    for mu in range(3):
+        for jj_mu in jj_vals[mu]:
+            op_list[jj_mu] = spin_op(mu)
+    return qt.tensor(op_list)
+
 def poch(nn, kk, vals = {}):
     return np.prod([ nn - cc for cc in range(kk) ])
 
@@ -60,49 +68,55 @@ def eta(mu, nu, rho, vals = {}):
         vals[mu,nu,rho] = ( spin_op(mu)*spin_op(nu) * spin_op(rho).dag() ).tr() / fac
         return vals[mu,nu,rho]
 
-def PP(jj,mm):
-    jj_vals = partition(jj,mm)
-    op_list = [ I2 ] * N
-    for mu in range(3):
-        for jj_mu in jj_vals[mu]:
-            op_list[jj_mu] = spin_op(mu)
-    return qt.tensor(op_list)
+def KK_cc(KK_mu_nu,cc,kk):
+    return KK_mu_nu[cc[:kk].sum():cc[:kk].sum()+cc[kk]]
 
-def KK_cc(KK,mu,nu,cc,kk):
-    return KK[mu][nu][cc[:kk].sum():cc[:kk].sum()+cc[kk]]
-
-def cc_vals(rr_mu_nu):
+def cc_rr_mu_nu(rr_mu_nu):
     return ( np.array([c_0,c_1,c_2,rr_mu_nu-c_0-c_1-c_2])
              for c_0 in range(rr_mu_nu+1)
              for c_1 in range(rr_mu_nu-c_0+1)
              for c_2 in range(rr_mu_nu-c_0-c_1+1) )
 
+def cc_rr(rr):
+    return [ [ cc_rr_mu_nu(rr[mu,nu]) for nu in range(3) ] for mu in range(3) ]
+
+def cc_ss(ss):
+    return ( np.array([c_0,c_1,c_2,ss-c_0-c_1-c_2])
+             for c_0 in range(ss+1)
+             for c_1 in range(ss-c_0+1)
+             for c_2 in range(ss-c_0-c_1+1) )
+
 def QQ(KK,rr):
-    KK_mat = partition(KK,rr)
+    QQ = 0 * II
+    ss = rr.sum()
+    for rho_vec in itertools.product(range(4), repeat = ss):
[…]

### 2018-10-02 soc_squeezing/cumulant_expansions.tex
more simplifying of the operator product, and code to test it

--- 2018-09-29 soc_squeezing/cumulant_expansions.tex
@@
 \end{align}
-where $\v m \equiv \p{m_+,m_\z,m_-}$; $\v j$ is an ordered list of
+where $\v m \equiv \p{m_\z,m_+,m_-}$; $\v j$ is an ordered list of
 indices for the spins addressed by $P_{\v j}$; and the sum over
@@
 indices in $\v j$ are distinct and $\abs{\set{j_a^\mu}} = m_\mu$.  In
-words, $\S_{\v m}$ is an exchange-symmetric $\p{m_++m_\z+m_-}$-spin
+words, $\S_{\v m}$ is an exchange-symmetric $\p{m_\z+m_++m_-}$-spin
 operator with $m_\mu$ single-spin operators of type $\sigma_\mu$, and
@@
   \S_{\v m}\S_{\v n}
-  = \sum_{s\ge0} \sum_{\p{\v r,\v J,\v K;\v m,\v n,s}}
-  g_{\v m\v n\v r} Q_{\v J} \tilde Q_{\v K},
+  = \sum_{s\ge0} \sum_{\p{\v J,\v K,\v r;\v m,\v n,s}}
+  f_{\v m\v n\v r} P_{\v J} Q_{\v K},
   \label{eq:SS_PQ}
@@
 the expansion for $\S_{\v m} \S_{\v n}$ in \eqref{eq:SS_PP}; the sum
-over $\p{\v r,\v J,\v K;\v m,\v n,s}$ implies a sum over all possible
-values of $\v r,\v J,\v K$ with the restrictions
+over $\p{\v J,\v K,\v r;\v m,\v n,s}$ denotes a sum over all values of
+$\v J,\v K,\v r$ with the restrictions
 \begin{align}
@@
   \abs{\set{J_a^\mu}}
-  = m_\mu + n_\mu - \sum_\rho\p{r_{\mu\rho}+r_{\rho\mu}},
-  &&
-  \abs{\set{K_a^{\mu\nu}}} = r_{\mu\nu},
-  &&
-  \set{J_a^\mu} \cap \set{K_a^{\mu\nu}} = \set{};
-\end{align}
-the factor $g_{\v m\v n\v r}$ counts the number of ways to pair
+  = m_\mu + n_\mu - \sum_\kappa\p{r_{\mu\kappa}+r_{\kappa\mu}},
+  &&
+  \abs{\set{K^{\mu\nu}_a}} = r_{\mu\nu},
+  &&
+  \set{J_a^\lambda} \cap \set{K_b^{\mu\nu}} = \set{};
+\end{align}
+the factor $f_{\v m\v n\v r}$ counts the number of ways to pair
 operators in $P_{\v j}$ and $P_{\v k}$ as prescribed by $\v r$, or
 \begin{align}
-  g_{\v m\v n\v r}
+  f_{\v m\v n\v r}
   &\equiv \sp{\prod_{\mu,\nu}
-    { m_\mu - \sum_{\rho<\nu} r_{\mu\rho} \choose r_{\mu\nu} }
-    { n_\mu - \sum_{\rho<\nu} r_{\rho\mu} \choose r_{\nu\mu} }}
+    { m_\mu - \sum_{\kappa<\nu} r_{\mu\kappa} \choose r_{\mu\nu} }
+    { n_\mu - \sum_{\kappa<\nu} r_{\kappa\mu} \choose r_{\nu\mu} }}
   \sp{\prod_{\mu,\nu} r_{\mu\nu}!} \\
-  &= \sp{\prod_\mu \p{m_\mu}_{\sum_\rho r_{\mu\rho}}
-    \p{n_\mu}_{\sum_\rho r_{\rho\mu}}}
+  &= \sp{\prod_\mu \p{m_\mu}_{\sum_\kappa r_{\mu\kappa}}
+    \p{n_\mu}_{\sum_\kappa r_{\kappa\mu}}}
   \sp{\prod_{\mu,\nu} r_{\mu\nu}!}^{-1},
@@
 where $\p{n}_k\equiv\prod_{j=0}^{k-1}\p{n-j}$ is a falling factorial;
-and finally
-\begin{align}
-  Q_{\v J} \equiv \prod_{\mu,a} \sigma_\mu^{(J_a^\mu)},
-  &&
-  \tilde Q_{\v K} \equiv \prod_{\mu,\nu,a}
-  \sigma_\mu^{(K_a^{\mu\nu})} \sigma_\nu^{(K_a^{\mu\nu})}
+and finally, similarly to the definition of $P_{\v j}$ in
+\eqref{eq:SS_PP} we have
+\begin{align}
+  Q_{\v K} \equiv \prod_{\mu,\nu,a}
+  \sigma_\mu^{(K^{\mu\nu}_a)} \sigma_\nu^{(K^{\mu\nu}_a)}.
+\end{align}
+Expanding the product
+$\sigma_\mu\sigma_\nu=\sum_\kappa\eta_{\mu\nu\rho}\sigma_\rho$, where
+unlike the sums over $\mu,\nu$ in most of the work above, the sum over
+$\rho$ here includes an index for the identity operator
+$\sigma_0\equiv\1$, we can expand and collect terms to get
+\begin{align}
+  Q_{\v K}
   = \prod_{\mu,\nu,a} \sum_\rho
-  \eta_{\mu\nu\rho} \sigma_\rho^{(K_a^{\mu\nu})},
-  \label{eq:Q_JK}
-\end{align}
-where $\eta_{\mu\nu\rho}$ is defined by the product
-$\sigma_\mu\sigma_\nu\equiv\sum_\rho\eta_{\mu\nu\rho}\sigma_\rho$.
-Note that unlike the sums over $\mu$ and $\nu$ in the work above, the
-sum over $\rho$ in \eqref{eq:Q_JK} now includes an index for the
-identity operator $\sigma_0\equiv\1$.
+  \eta_{\mu\nu\rho} \sigma_\rho^{(K^{\mu\nu}_a)}
+  = \sum_{\v\rho} \prod_{\mu,\nu,a} \eta_{\mu\nu\rho^{\mu\nu}_a}
+  \sigma_{\rho^{\mu\nu}_a}^{(K^{\mu\nu}_a)},
+\end{align}
+where we have distributed a product of sums into a sum of products.
+Collecting together operators of type $\sigma_\kappa$, we then have
+\begin{align}
+  Q_{\v K}
+  = \sum_{\v\rho} \sp{\prod_{\mu,\nu,\kappa}
+    \p{\eta_{\mu\nu\kappa}}^{\tilde\rho^{\mu\nu}_\kappa}}
+  \prod_{\kappa,a} \sigma_{\kappa}^{(K^{\v\rho\kappa}_a)},
+\end{align}
+where $\tilde\rho_\kappa^{\mu\nu}$ is the number of elements in
+$\v\rho^{\mu\nu} \equiv
+\p{\rho^{\mu\nu}_1,\rho^{\mu\nu}_2,\cdots,\rho^{\mu\nu}_{r_{\mu\nu}}}$
+equal to $\kappa$, and $K^{\v\rho\kappa}_a$ indexes the $a$-th
+operator of type $\sigma_\kappa$ for each choice of $\v\rho$, such
+that for any fixed $\v\rho$ we have
+$\set{K^{\v\rho\kappa}_a}=\set{K^{\mu\nu}_b}$.  Exploiting the
+exchange symmetry of $Q_{\v K}$ under the sum in \eqref{eq:SS_PQ}, we
+can group together all terms which are equal up to a permutation of
+indices in $\v\rho^{\mu\nu}$, i.e.~terms with equal values of
+$\tilde\rho_\kappa^{\mu\nu}$, to write
+\begin{align}
+  Q_{\v K}
+  \simeq \sum_{\tilde{\v\rho}}
+  g_{\tilde{\v\rho}} \sp{\prod_{\mu,\nu,\kappa}
+    \p{\eta_{\mu\nu\kappa}}^{\tilde\rho^{\mu\nu}_\kappa}}
+  \prod_{\kappa,a} \sigma_{\kappa}^{(K^{\tilde{\v\rho}\kappa}_a)}
+  = \sum_{\tilde{\v\rho}} g_{\tilde{\v\rho}}
+  \sp{\prod_{\mu,\nu,\kappa}
+    \p{\eta_{\mu\nu\kappa}}^{\tilde\rho^{\mu\nu}_\kappa}}
+  P_{\v K_{\tilde{\v\rho}}},
+\end{align}
+where $\simeq$ denotes equality up to a permutation of spin indices;
+the degeneracy factor for each term is
+\begin{align}
+  g_{\tilde{\v\rho}}
+  \equiv \prod_{\mu,\nu,\kappa}
+  { r_{\mu\nu} - \sum_{\lambda<\kappa} \tilde\rho^{\mu\nu}_\lambda
+    \choose \tilde\rho^{\mu\nu}_\kappa }
+  = \prod_{\mu,\nu}
+  \f{r_{\mu\nu}!}{\prod_\kappa\tilde\rho^{\mu\nu}_\kappa!};
+\end{align}
+as before $K^{\tilde{\v\rho}\kappa}_a$ indexes the $a$-th operator of
+type $\sigma_\kappa$ for each partitioning $\tilde{\v\rho}$, such that
+for any fixed $\tilde{\v\rho}$ we have
+$\set{K^{\tilde{\v\rho}\kappa}_a}=\set{K^{\mu\nu}_b}$; and
+$\v K_{\tilde{\v\rho}}$ is an ordered list of the indices in
+$\set{K^{\tilde{\v\rho}\kappa}_a}$.
+
+% \begin{align}
+%   \S_{\v m}\S_{\v n}
+%   = \sum_{s\ge0} \sum_{\p{\v J,\v K,\v r;\v m,\v n,s}}
+%   \sum_{\tilde{\v\rho}}
+%   f_{\v m\v n\v r} g_{\tilde{\v\rho}}
+%   \sp{\prod_{\mu,\nu,\kappa}
+%     \p{\eta_{\mu\nu\kappa}}^{\tilde\rho^{\mu\nu}_\kappa}}
+%   P_{\v J} P_{\v K_{\tilde{\v\rho}}}
+% \end{align}
 

### 2018-10-02 soc_squeezing/scripts/test_product.py

@@ -8,6 +8,7 @@ from math import factorial
 from scipy.special import binom
 
 N = 3
+op_cap = 2
 
 I2 = qt.qeye(2)
 sz = qt.sigmaz()
@@ -18,7 +19,7 @@ II = qt.tensor([ I2 ] * N)
 def partition_1D(indices, partition_sizes):
     return [ indices[partition_sizes[:mu].sum() :
                      partition_sizes[:mu].sum() + partition_sizes[mu] ]
-                     for mu in range(3) ]
+                     for mu in range(partition_sizes.size) ]
 
 def partition_2D(indices, partition_sizes):
     split_indices = partition_1D(indices, partition_sizes.sum(1))
@@ -41,10 +42,10 @@ def spin_op(mu):
     if mu == 2: return sm
     if mu == 3: return I2
 
-def PP(jj,mm):
+def PP(jj, mm):
     jj_vals = partition(jj,mm)
     op_list = [ I2 ] * N
-    for mu in range(3):
+    for mu in range(mm.size):
         for jj_mu in jj_vals[mu]:
             op_list[jj_mu] = spin_op(mu)
     return qt.tensor(op_list)
@@ -53,69 +54,44 @@ def poch(nn, kk, vals = {}):
     return np.prod([ nn - cc for cc in range(kk) ])
 
 def f(mm,nn,rr):
-    mnb_fac = np.prod([ poch(mm[mu],rr[mu,:].sum()) * poch(nn[mu],rr[:,mu].sum())
-                       for mu in range(3) ])
-    rr_fac = np.prod([ factorial(rr[mu,nu]) for mu in range(3) for nu in range(3) ])
-    return mnb_fac / rr_fac
+    return np.prod([ poch(mm[mu],rr[mu,:].sum()) * poch(nn[mu],rr[:,mu].sum())
+                     for mu in range(3) ])
 
-def g(cc,rr_mu_nu):
-    return factorial(rr_mu_nu) / np.prod([ factorial(cc_jj) for cc_jj in cc  ])
+def g(rho):
+    return 1 / np.prod([ factorial(val) for val in rho.flatten() ])
 
-def eta(mu, nu, rho, vals = {}):
-    try: return vals[mu,nu,rho]
-    except:
-        fac = 2 if rho in [1,3] else 1
-        vals[mu,nu,rho] = ( spin_op(mu)*spin_op(nu) * spin_op(rho).dag() ).tr() / fac
[…]

### 2018-10-02 soc_squeezing/cumulant_expansions.tex
greatly simplified a derivation

@@
   \abs{\set{J_a^\mu}}
-  = m_\mu + n_\mu - \sum_\kappa\p{r_{\mu\kappa}+r_{\kappa\mu}},
+  = m_\mu + n_\mu - \sum_\rho\p{r_{\mu\rho}+r_{\rho\mu}},
   &&
@@
   &\equiv \sp{\prod_{\mu,\nu}
-    { m_\mu - \sum_{\kappa<\nu} r_{\mu\kappa} \choose r_{\mu\nu} }
-    { n_\mu - \sum_{\kappa<\nu} r_{\kappa\mu} \choose r_{\nu\mu} }}
+    { m_\mu - \sum_{\rho<\nu} r_{\mu\rho} \choose r_{\mu\nu} }
+    { n_\mu - \sum_{\rho<\nu} r_{\rho\mu} \choose r_{\nu\mu} }}
   \sp{\prod_{\mu,\nu} r_{\mu\nu}!} \\
-  &= \sp{\prod_\mu \p{m_\mu}_{\sum_\kappa r_{\mu\kappa}}
-    \p{n_\mu}_{\sum_\kappa r_{\kappa\mu}}}
+  &= \sp{\prod_\mu \p{m_\mu}_{\sum_\rho r_{\mu\rho}}
+    \p{n_\mu}_{\sum_\rho r_{\rho\mu}}}
   \sp{\prod_{\mu,\nu} r_{\mu\nu}!}^{-1},
@@
 \end{align}
-Expanding the product
-$\sigma_\mu\sigma_\nu=\sum_\kappa\eta_{\mu\nu\rho}\sigma_\rho$, where
-unlike the sums over $\mu,\nu$ in most of the work above, the sum over
-$\rho$ here includes an index for the identity operator
-$\sigma_0\equiv\1$, we can expand and collect terms to get
+As the sum in \eqref{eq:SS_PQ} is invariant under permutation of the
+indices in $\v K$, we can safely neglect keeping track of these
+indices and simply write
 \begin{align}
   Q_{\v K}
-  = \prod_{\mu,\nu,a} \sum_\rho
-  \eta_{\mu\nu\rho} \sigma_\rho^{(K^{\mu\nu}_a)}
-  = \sum_{\v\rho} \prod_{\mu,\nu,a} \eta_{\mu\nu\rho^{\mu\nu}_a}
-  \sigma_{\rho^{\mu\nu}_a}^{(K^{\mu\nu}_a)},
-\end{align}
-where we have distributed a product of sums into a sum of products.
-Collecting together operators of type $\sigma_\kappa$, we then have
+  = \prod_{\mu,\nu} \prod_{a=1}^{r_{\mu\nu}}
+  \sigma_\mu^{(K^{\mu\nu}_a)} \sigma_\nu^{(K^{\mu\nu}_a)}
+  \simeq \bigotimes_{\mu,\nu} \bigotimes_{a=1}^{r_{\mu\nu}}
+  \sigma_\mu \sigma_\nu
+  = \bigotimes_{\mu,\nu} \bigotimes_{a=1}^{r_{\mu\nu}}
+  \sum_\rho \eta_{\mu\nu\rho} \sigma_\rho,
+\end{align}
+where $\simeq$ denotes equality up to a re-indexing of spins; we have
+introduced explicit dependence on
+$r_{\mu\nu}=\abs{\set{K^{\mu\nu}_a}}$ for brevity; and
+$\eta_{\mu\nu\rho}$ is defined by
+$\sigma_\mu\sigma_\nu=\sum_\rho\eta_{\mu\nu\rho}\sigma_\rho$.  Unlike
+the sums over $\mu,\nu$ in most of the work above, the sum over $\rho$
+here includes an index for the identity operator $\sigma_0\equiv\1$.
+Distributing the product of sums into a sum of products gives us
 \begin{align}
   Q_{\v K}
-  = \sum_{\v\rho} \sp{\prod_{\mu,\nu,\kappa}
-    \p{\eta_{\mu\nu\kappa}}^{\tilde\rho^{\mu\nu}_\kappa}}
-  \prod_{\kappa,a} \sigma_{\kappa}^{(K^{\v\rho\kappa}_a)},
-\end{align}
-where $\tilde\rho_\kappa^{\mu\nu}$ is the number of elements in
+  \simeq \sum_{\v\rho} \bigotimes_{\mu,\nu,a}
+  \eta_{\mu\nu\rho^{\mu\nu}_a} \sigma_{\rho^{\mu\nu}_a}
+  = \sum_{\v\rho} \sp{\prod_{\mu,\nu,a}
+    \p{\eta_{\mu\nu\rho^{\mu\nu}_a}}^{\rho^{\mu\nu}_a}}
+  \bigotimes_{\mu,\nu,a} \sigma_{\rho^{\mu\nu}_a}.
+  \label{eq:Q_K_rho}
+\end{align}
+Letting $\tilde\rho^{\mu\nu}_\kappa$ denote the number of elements in
 $\v\rho^{\mu\nu} \equiv
 \p{\rho^{\mu\nu}_1,\rho^{\mu\nu}_2,\cdots,\rho^{\mu\nu}_{r_{\mu\nu}}}$
-equal to $\kappa$, and $K^{\v\rho\kappa}_a$ indexes the $a$-th
-operator of type $\sigma_\kappa$ for each choice of $\v\rho$, such
-that for any fixed $\v\rho$ we have
-$\set{K^{\v\rho\kappa}_a}=\set{K^{\mu\nu}_b}$.  Exploiting the
-exchange symmetry of $Q_{\v K}$ under the sum in \eqref{eq:SS_PQ}, we
-can group together all terms which are equal up to a permutation of
-indices in $\v\rho^{\mu\nu}$, i.e.~terms with equal values of
-$\tilde\rho_\kappa^{\mu\nu}$, to write
+that are equal to $\kappa$, we observe that two terms in
+\eqref{eq:Q_K_rho} with, say, $\v\rho=\v\rho_1$ and $\v\rho=\v\rho_2$
+are equal up to a permutation of indices if
+$\tilde{\v\rho_1}=\tilde{\v\rho_2}$.  We therefore collect such terms
+together to write
 \begin{align}
   Q_{\v K}
-  \simeq \sum_{\tilde{\v\rho}}
-  g_{\tilde{\v\rho}} \sp{\prod_{\mu,\nu,\kappa}
-    \p{\eta_{\mu\nu\kappa}}^{\tilde\rho^{\mu\nu}_\kappa}}
-  \prod_{\kappa,a} \sigma_{\kappa}^{(K^{\tilde{\v\rho}\kappa}_a)}
-  = \sum_{\tilde{\v\rho}} g_{\tilde{\v\rho}}
-  \sp{\prod_{\mu,\nu,\kappa}
-    \p{\eta_{\mu\nu\kappa}}^{\tilde\rho^{\mu\nu}_\kappa}}
-  P_{\v K_{\tilde{\v\rho}}},
-\end{align}
-where $\simeq$ denotes equality up to a permutation of spin indices;
-the degeneracy factor for each term is
+  \simeq \sum_{\tilde{\v\rho}} g_{\tilde{\v\rho}} \v\eta_{\tilde{\v\rho}}
+  \bigotimes_{\kappa}
+  \sigma_\kappa^{\otimes\sum_{\mu,\nu}\tilde\rho^{\mu\nu}_\kappa},
+  &&
+  \v\eta_{\tilde{\v\rho}}
+  \equiv \prod_{\mu,\nu,\kappa}
+  \p{\eta_{\mu\nu\kappa}}^{\tilde\rho^{\mu\nu}_\kappa},
+\end{align}
+where
 \begin{align}
@@
   = \prod_{\mu,\nu}
-  \f{r_{\mu\nu}!}{\prod_\kappa\tilde\rho^{\mu\nu}_\kappa!};
-\end{align}
-as before $K^{\tilde{\v\rho}\kappa}_a$ indexes the $a$-th operator of
-type $\sigma_\kappa$ for each partitioning $\tilde{\v\rho}$, such that
-for any fixed $\tilde{\v\rho}$ we have
-$\set{K^{\tilde{\v\rho}\kappa}_a}=\set{K^{\mu\nu}_b}$; and
-$\v K_{\tilde{\v\rho}}$ is an ordered list of the indices in
-$\set{K^{\tilde{\v\rho}\kappa}_a}$.
-
-% \begin{align}
-%   \S_{\v m}\S_{\v n}
-%   = \sum_{s\ge0} \sum_{\p{\v J,\v K,\v r;\v m,\v n,s}}
-%   \sum_{\tilde{\v\rho}}
-%   f_{\v m\v n\v r} g_{\tilde{\v\rho}}
-%   \sp{\prod_{\mu,\nu,\kappa}
-%     \p{\eta_{\mu\nu\kappa}}^{\tilde\rho^{\mu\nu}_\kappa}}
-%   P_{\v J} P_{\v K_{\tilde{\v\rho}}}
-% \end{align}
+  \f{r_{\mu\nu}!}{\prod_\kappa\tilde\rho^{\mu\nu}_\kappa!}
+\end{align}
+counts the degeneracy of terms in \eqref{eq:Q_K_rho}, i.e.~the number
+of $\v\rho$ which are consistent with $\tilde{\v\rho}$.
 

### 2018-10-02 soc_squeezing/cumulant_expansions.tex
more work on the product of two operators

@@
   \sum_{\mu,\nu} r_{\mu\nu} = s,
+  \label{eq:rest_r}
 \end{align}
@@
   \sum_\rho \eta_{\mu\nu\rho} \sigma_\rho,
+  \label{eq:Q_K_eta}
 \end{align}
@@
     \p{\eta_{\mu\nu\rho^{\mu\nu}_a}}^{\rho^{\mu\nu}_a}}
-  \bigotimes_{\mu,\nu,a} \sigma_{\rho^{\mu\nu}_a}.
+  \bigotimes_{\mu,\nu,a} \sigma_{\rho^{\mu\nu}_a},
   \label{eq:Q_K_rho}
 \end{align}
-Letting $\tilde\rho^{\mu\nu}_\kappa$ denote the number of elements in
-$\v\rho^{\mu\nu} \equiv
-\p{\rho^{\mu\nu}_1,\rho^{\mu\nu}_2,\cdots,\rho^{\mu\nu}_{r_{\mu\nu}}}$
+where implicitly $\sum_a\rho^{\mu\nu}_a=r_{\mu\nu}$, as
+$\p{\mu,\nu,a}$ essentially index a factor in \eqref{eq:Q_K_eta}, and
+$\rho^{\mu\nu}_a$ indexes one of the terms in that factor.  Letting
+$\tilde\rho^{\mu\nu}_\kappa$ denote the number of elements in
+$\p{\rho^{\mu\nu}_1,\rho^{\mu\nu}_2,\cdots,\rho^{\mu\nu}_{r_{\mu\nu}}}$
 that are equal to $\kappa$, we observe that two terms in
@@
 are equal up to a permutation of indices if
-$\tilde{\v\rho_1}=\tilde{\v\rho_2}$.  We therefore collect such terms
-together to write
-\begin{align}
-  Q_{\v K}
-  \simeq \sum_{\tilde{\v\rho}} g_{\tilde{\v\rho}} \v\eta_{\tilde{\v\rho}}
-  \bigotimes_{\kappa}
-  \sigma_\kappa^{\otimes\sum_{\mu,\nu}\tilde\rho^{\mu\nu}_\kappa},
-  &&
-  \v\eta_{\tilde{\v\rho}}
-  \equiv \prod_{\mu,\nu,\kappa}
-  \p{\eta_{\mu\nu\kappa}}^{\tilde\rho^{\mu\nu}_\kappa},
-\end{align}
-where
+$\tilde{\v\rho_1}=\tilde{\v\rho_2}$.  The degeneracy of terms in
+\eqref{eq:Q_K_rho}, i.e.~the number of $\v\rho$ which are consistent
+with $\tilde{\v\rho}$, is
 \begin{align}
@@
   = \prod_{\mu,\nu}
-  \f{r_{\mu\nu}!}{\prod_\kappa\tilde\rho^{\mu\nu}_\kappa!}
-\end{align}
-counts the degeneracy of terms in \eqref{eq:Q_K_rho}, i.e.~the number
-of $\v\rho$ which are consistent with $\tilde{\v\rho}$.
+  \f{r_{\mu\nu}!}{\prod_\kappa\tilde\rho^{\mu\nu}_\kappa!}.
+\end{align}
+Defining additionally
+\begin{align}
+  \v\eta_{\tilde{\v\rho}}
+  \equiv \prod_{\mu,\nu,\kappa}
+  \p{\eta_{\mu\nu\kappa}}^{\tilde\rho^{\mu\nu}_\kappa},
+\end{align}
+we can collect like factors and group together equivalent terms in
+\eqref{eq:Q_K_rho} to write
+\begin{align}
+  Q_{\v K}
+  \simeq \sum_{\tilde{\v\rho}} g_{\tilde{\v\rho}} \v\eta_{\tilde{\v\rho}}
+  \bigotimes_{\kappa}
+  \sigma_\kappa^{\otimes\sum_{\mu,\nu}\tilde\rho^{\mu\nu}_\kappa}
+  = \sum_{\tilde{\v\rho}} g_{\tilde{\v\rho}} \v\eta_{\tilde{\v\rho}}
+  \1^{\otimes\sum_{\mu,\nu}\tilde\rho^{\mu\nu}_0}
+  \bigotimes_{\kappa\ne0}
+  \sigma_\kappa^{\otimes\sum_{\mu,\nu}\tilde\rho^{\mu\nu}_\kappa},
+  \label{eq:Q_K_rho_simp}
+\end{align}
+where again implicitly
+$\sum_\kappa\tilde\rho^{\mu\nu}_\kappa=r_{\mu\nu}$ for consistency
+with \eqref{eq:Q_K_eta} and \eqref{eq:Q_K_rho}, and we have explicitly
+factored out the identity operators in each term of $Q_{\v K}$.
+
+We have now simplified $Q_{\v K}$ sufficiently to substitute it back
+into \eqref{eq:SS_PQ}, which (removing the tilde from
+$\tilde{\v\rho}$) gives us
+\begin{align}
+  \S_{\v m} \S_{\v n}
+  = \sum_{s\ge0} \sum_{\p{\v\ell,\v r,\v\rho;\v m,\v n,s}}
+  f_{\v m\v n\v r\v\rho} P_{\v\ell},
+  \label{eq:SS_simp}
+\end{align}
+where the net the prefactor for each term is
+\begin{align}
+  f_{\v m\v n\v r\v\rho}
+  \equiv \sp{\prod_\mu \p{m_\mu}_{\sum_\kappa r_{\mu\kappa}}
+    \p{n_\mu}_{\sum_\kappa r_{\kappa\mu}}}
+  \sp{\prod_{\mu,\nu,\kappa}
+    \f{\p{\eta_{\mu\nu\kappa}}^{\rho^{\mu\nu}_\kappa}}
+    {\rho^{\mu\nu}_\kappa!}}
+  \sp{N-\sum_\mu\p{m_\mu+n_\mu}+s}
+  ^{\sum_{\alpha,\beta}\rho^{\alpha\beta}_0},
+\end{align}
+and in addition to \eqref{eq:rest_r} the sum over
+$\p{\v\ell,\v r,\v\rho;\v m,\v n,s}$ enforces the restrictions
+\begin{align}
+  \sum_{\kappa} \rho^{\mu\nu}_\kappa = r_{\mu\nu},
+  &&
+  \abs{\set{\ell^\mu_a}}
+  = m_\mu + n_\mu - \sum_\kappa \p{r_{\mu\kappa}+r_{\kappa\mu}}
+  + \sum_{\alpha,\beta} \rho^{\alpha\beta}_\mu.
+\end{align}
+For clarity, we note here that all indices on
+$m_\mu,n_\mu,r_{\mu\nu},\eta_{\mu\nu\kappa},\rho^{\mu\nu}_\kappa$ and
+the upper index on $\ell^\mu_a$ take on the same values as the indices
+on elements $\sigma_\mu$ of some complete basis for single-spin
+operators; of these, only the last index ($\kappa$) on
+$\eta_{\mu\nu\kappa},\rho^{\mu\nu}_\kappa$ ever matches the index for
+the identity element $\sigma_0\equiv\1$.  The lower index on
+$\ell^\mu_a$ indexes the $a$-th spin acted on by a $\sigma_\mu$
+operator in each term of \eqref{eq:SS_simp}.
 

### 2018-10-02 soc_squeezing/cumulant_expansions.tex
finished computing product of multi-body operators, as well as code to test it; now all that's left is simplifying the final result, if possible

@@
 We have now simplified $Q_{\v K}$ sufficiently to substitute it back
-into \eqref{eq:SS_PQ}, which (removing the tilde from
-$\tilde{\v\rho}$) gives us
+into \eqref{eq:SS_PQ}, which (removing the tilde from $\tilde{\v\rho}$
+to simplify notation) gives us
 \begin{align}
@@
   = \sum_{s\ge0} \sum_{\p{\v\ell,\v r,\v\rho;\v m,\v n,s}}
-  f_{\v m\v n\v r\v\rho} P_{\v\ell},
+  f_{\v m\v n\v r} g_{\v\rho} \v\eta_{\v\rho} c_{\v\ell\v\rho} P_{\v\ell},
   \label{eq:SS_simp}
 \end{align}
-where the net the prefactor for each term is
-\begin{align}
-  f_{\v m\v n\v r\v\rho}
-  \equiv \sp{\prod_\mu \p{m_\mu}_{\sum_\kappa r_{\mu\kappa}}
-    \p{n_\mu}_{\sum_\kappa r_{\kappa\mu}}}
-  \sp{\prod_{\mu,\nu,\kappa}
-    \f{\p{\eta_{\mu\nu\kappa}}^{\rho^{\mu\nu}_\kappa}}
-    {\rho^{\mu\nu}_\kappa!}}
-  \sp{N-\sum_\mu\p{m_\mu+n_\mu}+s}
-  ^{\sum_{\alpha,\beta}\rho^{\alpha\beta}_0},
-\end{align}
-and in addition to \eqref{eq:rest_r} the sum over
+where in addition to \eqref{eq:rest_r} the sum over
 $\p{\v\ell,\v r,\v\rho;\v m,\v n,s}$ enforces the restrictions
@@
   &&
-  \abs{\set{\ell^\mu_a}}
+  \abs{\v\ell} \equiv \abs{\set{\ell^\mu_a}}
   = m_\mu + n_\mu - \sum_\kappa \p{r_{\mu\kappa}+r_{\kappa\mu}}
-  + \sum_{\alpha,\beta} \rho^{\alpha\beta}_\mu.
-\end{align}
+  + \sum_{\alpha,\beta} \rho^{\alpha\beta}_\mu,
+\end{align}
+and the combinatorial factor
+\begin{align}
+  c_{\v\ell\v\rho}
+  \equiv { N - \abs{\v\ell} \choose \sum_{\mu,\nu}\rho^{\mu\nu}_0},
+\end{align}
+accounts for the multiplicity of terms which are equivalent up to a
+permutation of indices involving the identity operators in
+\eqref{eq:Q_K_rho_simp}.
+
 For clarity, we note here that all indices on

### 2018-10-02 soc_squeezing/scripts/test_product.py

@@ -50,15 +50,14 @@ def PP(jj, mm):
             op_list[jj_mu] = spin_op(mu)
     return qt.tensor(op_list)
 
+def SS(mm):
+    op = 0 * II
+    for jj in itertools.permutations(range(N), int(np.sum(mm))):
+        op += PP(jj,mm)
+    return op
+
 def poch(nn, kk, vals = {}):
     return np.prod([ nn - cc for cc in range(kk) ])
-
-def f(mm,nn,rr):
-    return np.prod([ poch(mm[mu],rr[mu,:].sum()) * poch(nn[mu],rr[:,mu].sum())
-                     for mu in range(3) ])
-
-def g(rho):
-    return 1 / np.prod([ factorial(val) for val in rho.flatten() ])
 
 def eta_val(mu, nu, rho):
     fac = 2 if spin_op(rho) in [ I2, sz ] else 1
@@ -71,15 +70,6 @@ nonzero = np.array([ [ [ eta[mu,nu,kk] != 0 for kk in range(4) ]
                        for nu in range(3) ]
                      for mu in range(3) ])
 eta_nonzero = eta[nonzero]
-
-def rho_vals(rr):
-    return np.array([ [ [ [0,0,0,rr[0,0]], [0,rr[0,1],0,0],     [0,0,rr[0,2],0],    ],
-                        [ [0,rr[1,0],0,0], [0,0,0,0],           [c_12_0,0,0,c_12_3] ],
-                        [ [0,0,rr[2,0],0], [c_21_0,0,0,c_21_3], [0,0,0,0]           ] ]
-                      for c_12_0 in range(rr[1,2]+1)
-                      for c_21_0 in range(rr[2,1]+1)
-                      for c_12_3 in [ rr[1,2] - c_12_0 ]
-                      for c_21_3 in [ rr[2,1] - c_21_0 ] ])
 
 def r_mats(mm,nn,ss):
     return ( np.array([ [r_00, r_01, r_02],
@@ -94,15 +84,14 @@ def r_mats(mm,nn,ss):
              for r_20 in [ ss-r_00-r_01-r_10-r_12-r_21-r_02 ]
              if r_20 + r_21 <= mm[2] and r_20+r_10+r_00 <= nn[0] )
 
-def QQ(KK,rr):
-    return np.sum( g(rho) * np.prod(eta_nonzero**rho[nonzero]) * PP(KK, rho.sum((0,1)))
[…]

### 2018-10-02 soc_squeezing/cumulant_expansions.tex
improved some notations

@@
   \S_{\v m}\S_{\v n}
-  = \sum_{s\ge0} \sum_{\p{\v J,\v K,\v r;\v m,\v n,s}}
-  f_{\v m\v n\v r} P_{\v J} Q_{\v K},
+  = \sum_{s\ge0} \sum_{\p{\v r;\v m,\v n,s}}
+  \sum_{\p{\v J,\v K;\v m,\v n,\v r}} f_{\v m\v n\v r} P_{\v J} Q_{\v K},
   \label{eq:SS_PQ}
@@
 the expansion for $\S_{\v m} \S_{\v n}$ in \eqref{eq:SS_PP}; the sum
-over $\p{\v J,\v K,\v r;\v m,\v n,s}$ denotes a sum over all values of
-$\v J,\v K,\v r$ with the restrictions
+over $\p{\v r;\v m,\v n,s}$ denotes a sum over all values of $\v r$ with the restrictions
 \begin{align}
@@
   &&
-  \sum_{\mu,\nu} r_{\mu\nu} = s,
+  \sum_{\mu,\nu} r_{\mu\nu} = s;
   \label{eq:rest_r}
 \end{align}
+the sum over $\p{\v J,\v K;\v m,\v n,\v r}$ denotes a sum over all
+values of $\v J,\v K$ with the restrictions
 \begin{align}
@@
   \1^{\otimes\sum_{\mu,\nu}\tilde\rho^{\mu\nu}_0}
-  \bigotimes_{\kappa\ne0}
+  \otimes \bigotimes_{\kappa\ne0}
   \sigma_\kappa^{\otimes\sum_{\mu,\nu}\tilde\rho^{\mu\nu}_\kappa},
@@
   \S_{\v m} \S_{\v n}
-  = \sum_{s\ge0} \sum_{\p{\v\ell,\v r,\v\rho;\v m,\v n,s}}
+  = \sum_{s\ge0} \sum_{\p{\v r;\v m,\v n,s}} \sum_{\p{\v\rho;\v r}}
+  \sum_{\p{\v\ell;\v m,\v n,\v r,\v\rho}}
   f_{\v m\v n\v r} g_{\v\rho} \v\eta_{\v\rho} c_{\v\ell\v\rho} P_{\v\ell},
@@
 \end{align}
-where in addition to \eqref{eq:rest_r} the sum over
-$\p{\v\ell,\v r,\v\rho;\v m,\v n,s}$ enforces the restrictions
+where the sums over $\p{\v\rho;\v r}$ and
+$\p{\v\ell;\v m,\v n,\v r,\v\rho}$ respectively denote sums over all
+values of $\v\rho$ and $\v\ell$ with the restrictions
 \begin{align}
@@
   &&
-  \abs{\v\ell} \equiv \abs{\set{\ell^\mu_a}}
+  \abs{\set{\ell^\mu_a}}
   = m_\mu + n_\mu - \sum_\kappa \p{r_{\mu\kappa}+r_{\kappa\mu}}
@@
 \end{align}
+with $\abs{\v\ell}$ equal to the number of elements in $\v\ell$,
 accounts for the multiplicity of terms which are equivalent up to a
@@
 For clarity, we note here that all indices on
-$m_\mu,n_\mu,r_{\mu\nu},\eta_{\mu\nu\kappa},\rho^{\mu\nu}_\kappa$ and
+$m_\mu,n_\mu,r_{\mu\nu},\rho^{\mu\nu}_\kappa,\eta_{\mu\nu\kappa}$ and
 the upper index on $\ell^\mu_a$ take on the same values as the indices
@@
 operators; of these, only the last index ($\kappa$) on
-$\eta_{\mu\nu\kappa},\rho^{\mu\nu}_\kappa$ ever matches the index for
+$\rho^{\mu\nu}_\kappa,\eta_{\mu\nu\kappa}$ ever matches the index for
 the identity element $\sigma_0\equiv\1$.  The lower index on

### 2018-10-02 soc_squeezing/cumulant_expansions.tex
added text about avoiding null terms in the product of multi-body opeators

@@
   \sum_{\substack{\p{\v j;\v m},\p{\v k;\v n} \\
-      \abs{\set{j_\alpha^\mu}\cap\set{k_\beta^\nu}}=s}} X,
+      \abs{\set{j_\alpha^\mu}\cap\set{k_\beta^\nu}}=s}} X.
   \label{eq:SS_PP}
 \end{align}
-where the sum over $s$ terminates after $s = \sum_\mu\p{m_\mu+n_\mu}$.
 Collecting terms in which $r_{\mu\nu}$ of the $\sigma_\mu$ operators
@@
 
+The sums in \eqref{eq:SS_simp} generally involve a large number of
+terms; in practice, many of these terms will be equal to zero.  We
+can, however, impose additional restrictions on these sums in such a
+way as to throw out terms which vanish, and keep those which do not.
+Remembering that $s$ counts the number of spins which are addressed by
+both operators $P_{\v j}$ and $P_{\v k}$ in \eqref{eq:SS_PP}, the
+first restriction we can impose comes from recognizing that for given
+$\S_{\v m},\S_{\v n}$, the overlap $s$ is bounded as
+\begin{align}
+  \max\set{0,\sum_\mu\p{m_\mu+n_\mu}-N}
+  \le s \le \min\set{N,\sum_\mu\p{m_\mu+n_\mu}}.
+\end{align}
+In addition, the group multiplication table $\eta_{\mu\nu\kappa}$ may
+generally contain zeros that we can preemptively avoid with
+restrictions on $\v r$ and $\v\rho$.  Recalling that $r_{\mu\nu}$
+counts the number of operators addressed by
+$\sigma_\mu\sigma_\nu=\sum_\kappa\eta_{\mu\nu\kappa}\sigma_\kappa$, if
+$\eta_{\mu\nu\kappa}=0$ for all $\kappa$,
+i.e.~$\sigma_\mu\sigma_\nu=0$, then any term with $r_{\mu\nu}>0$ will
+vanish.  Our second restriction on the sums in \eqref{eq:SS_simp} is
+therefore to fix $r_{\mu\nu}=r_{\nu\mu}=0$ for all $\mu,\nu$ for which
+$\sigma_\mu\sigma_\nu=0$.  Finally, we can avoid the nullifying effect
+of any additional zeros in the multiplication table
+$\eta_{\mu\nu\kappa}$ by restricting $\rho^{\mu\nu}_\kappa=0$ for all
+$\mu,\nu,\kappa$ for which $\eta_{\mu\nu\kappa}=0$.  These
+restrictions are sufficient to ensure that all individual terms in
+\eqref{eq:SS_simp} are nonzero, although they do not rule out the
+possibility of several terms canceling out.
+
 \end{document}

### 2018-10-03 soc_squeezing/cumulant_expansions.tex
minor wording changes

@@
   \p{\eta_{\mu\nu\kappa}}^{\tilde\rho^{\mu\nu}_\kappa},
+  \label{eq:eta_rho}
 \end{align}
@@
 \end{align}
-In addition, the group multiplication table $\eta_{\mu\nu\kappa}$ may
+In addition, the multiplication table $\eta_{\mu\nu\kappa}$ may
 generally contain zeros that we can preemptively avoid with
@@
 vanish.  Our second restriction on the sums in \eqref{eq:SS_simp} is
-therefore to fix $r_{\mu\nu}=r_{\nu\mu}=0$ for all $\mu,\nu$ for which
+therefore to fix $r_{\mu\nu}=r_{\nu\mu}=0$ for all $\mu,\nu$ with
 $\sigma_\mu\sigma_\nu=0$.  Finally, we can avoid the nullifying effect
 of any additional zeros in the multiplication table
-$\eta_{\mu\nu\kappa}$ by restricting $\rho^{\mu\nu}_\kappa=0$ for all
-$\mu,\nu,\kappa$ for which $\eta_{\mu\nu\kappa}=0$.  These
+$\eta_{\mu\nu\kappa}$, which will appear in $\v\eta_{\v\rho}$ as
+defined in \eqref{eq:eta_rho}, by restricting $\rho^{\mu\nu}_\kappa=0$
+for all $\mu,\nu,\kappa$ with $\eta_{\mu\nu\kappa}=0$.  These
 restrictions are sufficient to ensure that all individual terms in

### 2018-10-03 soc_squeezing/cumulant_expansions.tex
fixed error in combinatorial factor and modified some text

@@
   c_{\v\ell\v\rho}
-  \equiv { N - \abs{\v\ell} \choose \sum_{\mu,\nu}\rho^{\mu\nu}_0},
-\end{align}
-with $\abs{\v\ell}$ equal to the number of elements in $\v\ell$,
-accounts for the multiplicity of terms which are equivalent up to a
-permutation of indices involving the identity operators in
-\eqref{eq:Q_K_rho_simp}.
+  \equiv \p{ N - \abs{\v\ell} }_ {\sum_{\mu,\nu}\rho^{\mu\nu}_0},
+\end{align}
+with $\abs{\v\ell}=\sum_\mu\p{m_\mu+n_\mu}-s$ equal to the number of
+elements in $\v\ell$, accounts for the multiplicity of terms which are
+equivalent up to a permutation of indices involving the identity
+operators in \eqref{eq:Q_K_rho_simp}.
 
@@
 of any additional zeros in the multiplication table
-$\eta_{\mu\nu\kappa}$, which will appear in $\v\eta_{\v\rho}$ as
-defined in \eqref{eq:eta_rho}, by restricting $\rho^{\mu\nu}_\kappa=0$
-for all $\mu,\nu,\kappa$ with $\eta_{\mu\nu\kappa}=0$.  These
-restrictions are sufficient to ensure that all individual terms in
-\eqref{eq:SS_simp} are nonzero, although they do not rule out the
-possibility of several terms canceling out.
+$\eta_{\mu\nu\kappa}$, which appear as factors in
+$\v\eta_{\v\rho}\propto\p{\eta_{\mu\nu\kappa}}^{\rho^{\mu\nu}_\kappa}$
+as defined in \eqref{eq:eta_rho}, by restricting
+$\rho^{\mu\nu}_\kappa=0$ for all $\mu,\nu,\kappa$ with
+$\eta_{\mu\nu\kappa}=0$.  These restrictions are sufficient to ensure
+that all individual terms in \eqref{eq:SS_simp} are nonzero, although
+they do not rule out the possibility of several terms canceling out.
 

### 2018-10-03 soc_squeezing/cumulant_expansions.tex
realized that calculation of the product of two operators was not actually completed... it now is

@@
   \sum_{\p{\v\ell;\v m,\v n,\v r,\v\rho}}
-  f_{\v m\v n\v r} g_{\v\rho} \v\eta_{\v\rho} c_{\v\ell\v\rho} P_{\v\ell},
+  f_{\v m\v n\v r} g_{\v\rho} \v\eta_{\v\rho} c_{\v\ell\v\rho} \S_{\v\ell},
   \label{eq:SS_simp}
@@
   &&
-  \abs{\set{\ell^\mu_a}}
+  \ell_\mu
   = m_\mu + n_\mu - \sum_\kappa \p{r_{\mu\kappa}+r_{\kappa\mu}}
@@
   c_{\v\ell\v\rho}
-  \equiv \p{ N - \abs{\v\ell} }_ {\sum_{\mu,\nu}\rho^{\mu\nu}_0},
-\end{align}
-with $\abs{\v\ell}=\sum_\mu\p{m_\mu+n_\mu}-s$ equal to the number of
-elements in $\v\ell$, accounts for the multiplicity of terms which are
-equivalent up to a permutation of indices involving the identity
-operators in \eqref{eq:Q_K_rho_simp}.
-
-For clarity, we note here that all indices on
-$m_\mu,n_\mu,r_{\mu\nu},\rho^{\mu\nu}_\kappa,\eta_{\mu\nu\kappa}$ and
-the upper index on $\ell^\mu_a$ take on the same values as the indices
-on elements $\sigma_\mu$ of some complete basis for single-spin
-operators; of these, only the last index ($\kappa$) on
+  \equiv \p{ N - \sum_\kappa\ell_\kappa }_{\sum_{\mu,\nu}\rho^{\mu\nu}_0},
+\end{align}
+accounts for the multiplicity of terms which are equivalent up to a
+permutation of indices involving the identity operators in
+\eqref{eq:Q_K_rho_simp}.  For clarity, we note here that all indices
+on
+$m_\mu, n_\mu, r_{\mu\nu}, \rho^{\mu\nu}_\kappa, \eta_{\mu\nu\kappa},
+\ell_\mu$ take on the same values as the indices on elements
+$\sigma_\mu$ of some complete basis for single-spin operators; of
+these, only the last index ($\kappa$) on
 $\rho^{\mu\nu}_\kappa,\eta_{\mu\nu\kappa}$ ever matches the index for
-the identity element $\sigma_0\equiv\1$.  The lower index on
-$\ell^\mu_a$ indexes the $a$-th spin acted on by a $\sigma_\mu$
-operator in each term of \eqref{eq:SS_simp}.
+the identity element $\sigma_0\equiv\1$.
 

### 2018-10-03 soc_squeezing/scripts/test_product.py

@@ -128,8 +128,7 @@ for mm in itertools.product(range(op_cap+1), repeat = 3):
                     id_fac = poch(N-ops,id_ops)
                     fac = mnr_fac * rho_fac * eta_fac * id_fac
 
-                    for ll in itertools.permutations(range(N), ops):
-                        op_simp += fac * PP(ll, op_nums)
+                    op_simp += fac * SS(op_nums)
 
         print(mm, nn, op == op_simp)
         if op != op_simp: exit()

### 2018-10-04 soc_squeezing/cumulant_expansions.tex
rearranged some text in a derivation for clarity

@@
   \S_{\v m}\S_{\v n}
-  = \sum_{s\ge0} \sum_{\p{\v r;\v m,\v n,s}}
-  \sum_{\p{\v J,\v K;\v m,\v n,\v r}} f_{\v m\v n\v r} P_{\v J} Q_{\v K},
+  = \sum_{s\ge0} \sum_{\p{\v r;\v m,\v n,s}} f_{\v m\v n\v r}
+  \sum_{\p{\v J,\v K;\v m,\v n,\v r}} P_{\v J} Q_{\v K},
   \label{eq:SS_PQ}
 \end{align}
-where $\v J$ and $\v K$ are respectively ordered lists of indices for
-spins addressed by one and two single-spin operators in each term of
-the expansion for $\S_{\v m} \S_{\v n}$ in \eqref{eq:SS_PP}; the sum
-over $\p{\v r;\v m,\v n,s}$ denotes a sum over all values of $\v r$ with the restrictions
+where the sum over $\p{\v r;\v m,\v n,s}$ denotes a sum over all
+values of $\v r$ with the restrictions
 \begin{align}
@@
 \end{align}
-the sum over $\p{\v J,\v K;\v m,\v n,\v r}$ denotes a sum over all
-values of $\v J,\v K$ with the restrictions
-\begin{align}
-  \abs{\set{J_a^\mu}}
-  = m_\mu + n_\mu - \sum_\rho\p{r_{\mu\rho}+r_{\rho\mu}},
-  &&
-  \abs{\set{K^{\mu\nu}_a}} = r_{\mu\nu},
-  &&
-  \set{J_a^\lambda} \cap \set{K_b^{\mu\nu}} = \set{};
-\end{align}
 the factor $f_{\v m\v n\v r}$ counts the number of ways to pair
-operators in $P_{\v j}$ and $P_{\v k}$ as prescribed by $\v r$, or
+operators in $P_{\v j}$ and $P_{\v k}$ according to $\v r$, or
 \begin{align}
@@
 \end{align}
-where $\p{n}_k\equiv\prod_{j=0}^{k-1}\p{n-j}$ is a falling factorial;
-and finally, similarly to the definition of $P_{\v j}$ in
-\eqref{eq:SS_PP} we have
+with $\p{n}_k\equiv\prod_{j=0}^{k-1}\p{n-j}$ denoting a falling
+factorial; $\v J$ and $\v K$ are respectively ordered lists of indices
+for spins addressed by one and two single-spin operators in a term of
+the product $P_{\v j} P_{\v k}$ with fixed $\v r$; the sum over
+$\p{\v J,\v K;\v m,\v n,\v r}$ denotes a sum over all values of
+$\v J,\v K$ with the restrictions
+\begin{align}
+  \abs{\set{J_a^\mu}}
+  = m_\mu + n_\mu - \sum_\rho\p{r_{\mu\rho}+r_{\rho\mu}},
+  &&
+  \abs{\set{K^{\mu\nu}_a}} = r_{\mu\nu},
+  &&
+  \set{J_a^\lambda} \cap \set{K_b^{\mu\nu}} = \set{};
+\end{align}
+and finally, similarly to $P_{\v j}$ in \eqref{eq:SS_PP} we define
 \begin{align}
@@
 \end{align}
-where implicitly $\sum_a\rho^{\mu\nu}_a=r_{\mu\nu}$, as
-$\p{\mu,\nu,a}$ essentially index a factor in \eqref{eq:Q_K_eta}, and
+where implicitly $\abs{\set{\rho^{\mu\nu}_a}}=r_{\mu\nu}$, as
+$\p{\mu,\nu,a}$ essentially index a factor in \eqref{eq:Q_K_eta} and
 $\rho^{\mu\nu}_a$ indexes one of the terms in that factor.  Letting
@@
   \S_{\v m} \S_{\v n}
-  = \sum_{s\ge0} \sum_{\p{\v r;\v m,\v n,s}} \sum_{\p{\v\rho;\v r}}
+  = \sum_{s\ge0} \sum_{\p{\v r;\v m,\v n,s}} f_{\v m\v n\v r}
+  \sum_{\p{\v\rho;\v r}} g_{\v\rho} \v\eta_{\v\rho}
   \sum_{\p{\v\ell;\v m,\v n,\v r,\v\rho}}
-  f_{\v m\v n\v r} g_{\v\rho} \v\eta_{\v\rho} c_{\v\ell\v\rho} \S_{\v\ell},
+  c_{\v\ell\v\rho} \S_{\v\ell},
   \label{eq:SS_simp}
@@
   c_{\v\ell\v\rho}
-  \equiv \p{ N - \sum_\kappa\ell_\kappa }_{\sum_{\mu,\nu}\rho^{\mu\nu}_0},
+  \equiv \p{ N - \sum_\kappa\ell_\kappa }_{\sum_{\mu,\nu}\rho^{\mu\nu}_0}
 \end{align}

### 2018-10-04 soc_squeezing/cumulant_expansions.tex
fixed error in operator product result

@@
 We wish to change our basis for all exchange-symmetric operators on a
-system of $N$ qubits from $S_+^\ell S_\z^m S_-^n$ to
+system of $N$ spins from $S_+^\ell S_\z^m S_-^n$ to
 \begin{align}
@@
   &&
-  P_{\v j} \equiv \prod_{\mu,a} \sigma_\mu^{\p{j_a^\mu}},
-\end{align}
-where $\v m \equiv \p{m_\z,m_+,m_-}$; $\v j$ is an ordered list of
-indices for the spins addressed by $P_{\v j}$; and the sum over
-$\p{\v j;\v m}$ denotes a sum over all possible $\v j$ for which all
-indices in $\v j$ are distinct and $\abs{\set{j_a^\mu}} = m_\mu$.  In
-words, $\S_{\v m}$ is an exchange-symmetric $\p{m_\z+m_++m_-}$-spin
-operator with $m_\mu$ single-spin operators of type $\sigma_\mu$, and
-$P_{\v j}$ is a single term in $\S_{\v m}$ which addresses the spins
-indexed by $\v j$.
+  P_{\v j} \equiv \prod_{\mu,a} \sigma_\mu^{\p{j^\mu_a}},
+\end{align}
+where $\v m \equiv \p{m_\z,m_+,m_-}$ with $m_\mu$ the number of
+$\sigma_\mu$ operators in each term of $\S_{\v m}$; $\v j$ is an
+ordered list of indices for the spins addressed by $P_{\v j}$; and the
+sum over $\p{\v j;\v m}$ denotes a sum over all possible $\v j$ for
+which all indices in $\v j$ are distinct and
+$\abs{\set{j^\mu_a}} = m_\mu$, such that $j^\mu_a$ indexes the $a$-th
+spin addressed by a $\sigma_\mu$ operator in $P_{\v j}$.
 
@@
   \sum_{\p{\v\rho;\v r}} g_{\v\rho} \v\eta_{\v\rho}
-  \sum_{\p{\v\ell;\v m,\v n,\v r,\v\rho}}
-  c_{\v\ell\v\rho} \S_{\v\ell},
+  c_{\v\ell_{\v m\v n\v r\v\rho}\v\rho} \S_{\v\ell_{\v m\v n\v r\v\rho}},
   \label{eq:SS_simp}
 \end{align}
-where the sums over $\p{\v\rho;\v r}$ and
-$\p{\v\ell;\v m,\v n,\v r,\v\rho}$ respectively denote sums over all
-values of $\v\rho$ and $\v\ell$ with the restrictions
-\begin{align}
-  \sum_{\kappa} \rho^{\mu\nu}_\kappa = r_{\mu\nu},
-  &&
-  \ell_\mu
+where the sum over $\p{\v\rho;\v r}$ denotes a sum over all
+values of $\v\rho$ and with the restrictions
+\begin{align}
+  \abs{\set{\rho^{\mu\nu}_\kappa}} = \abs{\set{\sigma_\kappa}},
+  &&
+  \sum_{\kappa} \rho^{\mu\nu}_\kappa = r_{\mu\nu};
+\end{align}
+the number of $\sigma_\mu$ operators in each term of
+$\S_{\v\ell_{\v m\v n\v r\v\rho}}$ is
+\begin{align}
+  \ell_{\v m\v n\v r\v\rho;\mu}
   = m_\mu + n_\mu - \sum_\kappa \p{r_{\mu\kappa}+r_{\kappa\mu}}
-  + \sum_{\alpha,\beta} \rho^{\alpha\beta}_\mu,
+  + \sum_{\alpha,\beta} \rho^{\alpha\beta}_\mu;
 \end{align}
@@
 \end{align}
-accounts for the multiplicity of terms which are equivalent up to a
+accounts for the number of terms which are equivalent up to a
 permutation of indices involving the identity operators in
-\eqref{eq:Q_K_rho_simp}.  For clarity, we note here that all indices
-on
-$m_\mu, n_\mu, r_{\mu\nu}, \rho^{\mu\nu}_\kappa, \eta_{\mu\nu\kappa},
-\ell_\mu$ take on the same values as the indices on elements
-$\sigma_\mu$ of some complete basis for single-spin operators; of
-these, only the last index ($\kappa$) on
-$\rho^{\mu\nu}_\kappa,\eta_{\mu\nu\kappa}$ ever matches the index for
-the identity element $\sigma_0\equiv\1$.
+\eqref{eq:Q_K_rho_simp}.
 

### 2018-10-13 soc_squeezing/cumulant_expansions.tex
made text in the main body more general, i.e. apply for any basis for collective spin operators

@@
 
-\title{Solving Collective Spin Hamiltonians with Decoherence}
+\title{Short-time simulation of collective spin Hamiltonians with
+  arbitrary decoherence}
 
@@
 
-Past work has shown how to realize spin-squeezing Hamiltonians with a
-variety of ultracold atomic and trapped ion systems.  While simulating
-the dynamics and squeezing behavior induced by these Hamiltonians is
-straightforward in the context of a closed quantum system, here we
-develop a method to simulate dynamics in the presence of decoherence.
-In particular, we will consider dynamics induced by the one-axis
-twisting (OAT), two-axis twisting (TAT), and transverse field (TVF)
-Hamiltonians of the form
+Previous work has shown how to realize spin-squeezing Hamiltonians
+with a variety of ultracold atomic and trapped ion systems.  While
+simulating the dynamics of initially Gaussian states evolving under
+these Hamiltonians is straightforward in the context of a closed
+quantum system, here we develop a method to simulate short-time
+dynamics in the presence of decoherence.  In particular, we will
+consider dynamics induced by the one-axis twisting (OAT), two-axis
+twisting (TAT), and transverse field (TVF) Hamiltonians of the form
 \begin{align}
@@
 where $\J$ is a set of jump operators with a corresponding decoherence
-rate $\gamma_\J$ and $\sp{A,B}_\pm\equiv AB\pm BA$.  Our measure of
+rate $\gamma_\J$ and $\sp{X,Y}_\pm\equiv XY\pm YX$.  Our measure of
 spin squeezing is
@@
 
-The case of the one-axis twisting is simplest to consider, as we can
-write the Hamiltonian in the form
+The simplest case is that of one-axis twisting, as we can write the
+Hamiltonian in the form
 \begin{align}
@@
 \end{align}
-defined respectively in terms of spontaneous uncorrelated spin
-dephasing, excitation, and decay rates $\gamma_\z$, $\gamma_+$,
-$\gamma_-$.  The last ingredients we need to compute spin squeezing at
-any time are $\bk{S_\z}$ and $\bk{S_\z^2}$.  As these operators
-commute with the OAT Hamiltonian, their evolution is governed entirely
-by decoherence (see Appendix \ref{sec:decay_single}), which means
+defined respectively in terms of uncorrelated spin dephasing,
+excitation, and decay rates $\gamma_\z$, $\gamma_+$, $\gamma_-$.  The
+last ingredients we need to compute spin squeezing at any time are
+$\bk{S_\z}$ and $\bk{S_\z^2}$.  As these operators commute with the
+OAT Hamiltonian, their evolution is governed entirely by decoherence
+(see Appendix \ref{sec:decay_single}), which means
 \begin{align}
@@
 $\xi^2$ defined in \eqref{eq:squeezing} at any time throughout
-evolution of the initial state $\ket{\X}:S_\x\ket{\X}=S\ket{\X}$ under
-$H_{\t{OAT}}$.
-
-\section{Truncated operator dynamics}
+evolution of the initial state $\ket{\X}$ under $H_{\t{OAT}}$.
+
+\section{Short-time evolution via truncated equations of motion}
 
 Unlike the one-axis twisting Hamiltonian, the two-axis twisting and
-transverse-field Hamiltonians do not admit exact analytical
-solutions. We will therefore solve these Hamiltonians approximately
-using alternate methods.  The basic strategy will be to find equations
-of motion for operators of interest, and at some point truncate
-high-order terms with negligible contribution to spin correlators of
-interest.  To benchmark this method, we test it against exact one-axis
-twisting results in the presence of decoherence, collective
-Hamiltonian dynamics in the absence of decoherence, and exact
-simulations of few-spin systems.
-
-All collective spin Hamiltonians can be written in the form
-\begin{align}
-  H = \sum_{\ell,m,n} h_{\ell mn} S_+^\ell S_\z^m S_-^n,
+transverse-field Hamiltonians do not admit exact analytical solutions.
+We will therefore simulate dynamics under these Hamiltonians
+numerically using alternate methods.  The basic strategy will be to
+first find equations of motion for all operators of interest using
+some complete basis for collective-spin operators.  These equations of
+motion will generally yield a system of equations involving a number
+of operators that grows too fast with system size for exact
+simulation.  We will therefore look for ways to truncate the equations
+of motion to arrive at a reasonably-sized expressions that are
+practical to compute.  To benchmark this method, we test it against
+collective Hamiltonian dynamics in the absence of decoherence, exact
+one-axis twisting results in the presence of decoherence, and exact
+simulations of small systems.
+
+Ultimately, we are interested in the dynamics of collective spin
+operators.  Choosing a basis $\set{\S_{\v m}}$ for these operators,
+e.g.~$\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$ with
+$\v m\equiv\p{m_+,m_\z,m_-}\in\mathbb{N}_0^3$, we can expand all
+collective-spin Hamiltonians in the form
+\begin{align}
+  H = \sum_{\v m} h_{\v m} \S_{\v m},
   \label{eq:H_general}
 \end{align}
-where Hermiticity requires that $h_{\ell mn}=h_{nm\ell}^*$.  The
-Hamiltonian for one-axis twisting with a transverse field, for
-example, is
+where if $\S_{\v m}=S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$ then Hermiticity
+requires that $h_{m_+,m_\z,m_-}=h_{m_-,m_\z,m_+}^*$.  The one-axis
+twisting Hamiltonian with a transverse field, for example, is
 \begin{align}
@@
 \end{align}
-while the two-axis twisting Hamiltonians about $\p{\z,\y}$ and
-$\p{\y,\x}$ are
+while the two-axis twisting Hamiltonians about $\p{z,y}$ and $\p{y,x}$
+are
 \begin{align}
@@
 \end{align}
-The evolution of a general operator $S_+^\ell S_\z^m S_-^n$ under a
-Hamiltonian of the form in \eqref{eq:H_general} is given by
-\begin{align}
-  \f{d}{dt} \p{S_+^\ell S_\z^m S_-^n}
-  = i \sum_{p,q,r} h_{pqr}
-  \sp{S_+^p S_\z^q S_-^r, S_+^\ell S_\z^m S_-^n}_-
-  + \sum_\J \gamma_\J \D\p{\J; S_+^\ell S_\z^m S_-^n}
-  \equiv \sum_{p,q,r} T^{\ell m n}_{pqr} S_+^p S_\z^q S_-^r,
+The evolution of a general operator $\S_{\v m}$ under a Hamiltonian of
+the form in \eqref{eq:H_general} is given by
+\begin{align}
+  \f{d}{dt} \S_{\v m}
+  = i \sum_{\v n} h_{\v n}
+  \sp{\S_{\v n}, \S_{\v m}}_-
+  + \sum_\J \gamma_\J \D\p{\J; \S_{\v m}}
+  \equiv \sum_{\v n} T_{\v m\v n} \S_{\v n},
   \label{eq:time_derivative}
@@
 where commutator and decoherence terms necessary for computing the
-matrix elements $T^{\ell m n}_{pqr}$ of the time derivative operator
-$d/dt$ are worked out Appendices \ref{sec:comm_general},
+matrix elements $T_{\v m\v n}$ of the time derivative operator $d/dt$
+are worked out Appendices \ref{sec:comm_general},
 \ref{sec:decoherence_single}, and \ref{sec:decoherence_collective}.
@@
 The time derivative operator $d/dt$ will generally couple spin
-operators $S_+^\ell S_\z^m S_-^n$ to spin operators
-$S_+^p S_\z^q S_-^r$ with higher ``weight'', i.e.~with
-$p+q+r>\ell+m+n$, which signifies the growth of many-body
-correlations.  In practice, keeping track of this operator growth will
-eventually require more computational resources than are available, so
-we must truncate it at some point by taking
-\begin{align}
-  \f{d}{dt} \p{S_+^\ell S_\z^m S_-^n}
-  = \sum_{p,q,r} T^{\ell m n}_{pqr} S_+^p S_\z^q S_-^r
-  \to \sum_{w\p{p,q,r}<W} T^{\ell m n}_{pqr} S_+^p S_\z^q S_-^r
+operators $\S_{\v m}$ to spin operators $\S_{\v n}$ with higher
+``weight'', i.e.~with $\abs{\v n}>\abs{\v m}$ where
+$\abs{\v\ell}\equiv\sum_\lambda\ell_\lambda$.  The growth of operator
+weight signifies the growth of many-body correlations.  In practice,
+keeping track of this growth will eventually require more
+computational resources than are available, so we truncate it by
+taking
+\begin{align}
+  \f{d}{dt} \S_{\v m}
+  = \sum_{\v n} T_{\v m\v n} \S_{\v n}
+  \to \sum_{w\p{\v n}<W} T_{\v m\v n} \S_{\v n}
   \label{eq:time_derivative_truncation}
 \end{align}
-for some cutoff $W$ and weight function $w$, e.g.~$w\p{p,q,r}=p+q+r$.
-The truncation in \eqref{eq:time_derivative_truncation} will close the
-system of differential equations defined by
+for some weight measure $w$, e.g.~$w\p{\v n}=\abs{\v n}$, and cutoff
+$W$.  The truncation in \eqref{eq:time_derivative_truncation} will
+close the system of differential equations defined by
 \eqref{eq:time_derivative}, and allow us to solve it using standard
@@
   of operators can be very time-consuming,
-\item the weight function $w$ may need to be chosen carefully, as the
-  optimal weight function is generally system-dependent,
+\item the weight measure $w$ may need to be chosen carefully, as the
+  optimal measure is generally system-dependent,
 \item simulation results can only be trusted up to the time at which
-  the initial value of operators $S_+^p S_\z^r S_-^q$ with weight
-  $w\p{p,q,r}\ge W$ have a non-negligible contribution to operators of
+  the initial value of operators $\S_{\v m}$ with weight
+  $w\p{\v m}\ge W$ have a non-negligible contribution to operators of
   interest.\label{pt:limitation}
 \end{enumerate*}
-While the first two of these limitations might in principle be avoided
-by alternate means of tracking operators, the last limitation
-necessarily applies in some form to any method tracking only a subset
-of {\it all} operators.  Embracing the unavoidable, we therefore try
-to devise an alternate truncation strategy which is built around
-\ref{pt:limitation}.
-
-We can expand operators in a Taylor series about $t=0$ to write
-\begin{align}
-  S_+^\ell S_\z^m S_-^n
+The last limitation in particular {\it necessarily} applies in some
+form to any method tracking only a subset of all relevant operators.
+Embracing the unavoidable, we therefore try to devise an alternate
+truncation strategy built around \ref{pt:limitation}.
+
+We can expand Heisenberg operators in a Taylor series about $t=0$ to
+write
+\begin{align}
+  \bk{\S_{\v m}}
+  = \sum_{k\ge0} \f{t^k}{k!} \bk{\f{d^k}{dt^k} \S_{\v m}}_{t=0}
   = \sum_{k\ge0} \f{t^k}{k!}
-  \left.\f{d^k}{dt^k}\right|_{t=0} \p{S_+^\ell S_\z^m S_-^n},
-\end{align}
-which implies
-\begin{align}
-  \bk{S_+^\ell S_\z^m S_-^n}
-  = \sum_{k\ge0} \f{t^k}{k!}
-  \sum_{p,q,r} T^{\ell m n}_{pqr;k} \bk{S_+^p S_\z^q S_-^r}_{t=0},
+  \sum_{\v n} T_{\v m\v n;k} \bk{\S_{\v n}}_{t=0},
   \label{eq:time_series}
@@
 \begin{align}
-  T^{\ell m n}_{pqr;0}=\delta_{\ell p}\delta_{mq}\delta_{nr},
-  &&
-  T^{\ell m n}_{pqr;1}\equiv T^{\ell m n}_{pqr},
-\end{align}
-\begin{align}
-  T^{\ell mn}_{pqr;k>1}
-  \equiv \sum_{\v\lambda,\v\mu,\v\nu}
-  T^{\ell mn}_{\lambda_1 \mu_1 \nu_1}
-  T^{\lambda_1 \mu_1 \nu_1}_{\lambda_2 \mu_2 \nu_2}
-  T^{\lambda_2 \mu_2 \nu_2}_{\lambda_3 \mu_3 \nu_3}
-  \cdots T^{\lambda_{k-1} \mu_{k-1} \nu_{k-1}}_{pqr}.
+  T_{\v m\v n;0} \equiv \delta_{\v m\v n},
+  &&
+  T_{\v m\v n;1} \equiv T_{\v m\v n},
+  &&
+  T_{\v m\v n;k>1}
+  \equiv \sum_{\v p_1,\v p_2,\cdots\v p_{k-1}}
+  T_{\v m\v p_1} T_{\v p_1\v p_2} T_{\v p_2\v p_3}
+  \cdots T_{\v p_{k-1}\v n}.
 \end{align}
@@
 \begin{align}
-  \bk{S_+^\ell S_\z^m S_-^n}
+  \bk{\S_{\v m}}
   = \sum_{k\ge0} \f{t^k}{k!}
-  \sum_{p,q,r} T^{\ell m n}_{pqr;k} \bk{S_+^p S_\z^q S_-^r}_{t=0}
-  \to \sum_{k<M} \f{t^k}{k!}
-  \sum_{p,q,r} T^{\ell m n}_{pqr;k} \bk{S_+^p S_\z^q S_-^r}_{t=0}.
-\end{align}
-Examining the nonzero entries of $T^{\ell mn}_{pqr;k}$ for
-$k=0,1,\cdots,M-1$ now {\it tells} us which operators we need to keep
-track of in order to compute the expectation value
-$\bk{S_+^\ell S_\z^m S_-^n}$.  Furthermore, after computing all
-relevant matrix elements $T^{\ell mn}_{pqr;k}$ and expectation values
-$\bk{S_+^p S_\z^q S_-^r}_{t=0}$, computing
-$\bk{S_+^\ell S_\z^m S_-^n}$ takes only a short time independent of
-system size.  Finally, the relation
-$\p{S_+^\ell S_\z^m S_-^n}^\dag=S_+^n S_\z^m S_-^\ell$ can be
-exploited to cut both the number of matrix elements
-$T^{\ell mn}_{pqr}$ and expectation values
-$\bk{S_+^p S_\z^r S_-^q}_{t=0}$ that we need to compute roughly in
-half.
-
-\section{Decoherence in the two-axis twisting protocols}
-
-In the case of squeezing through spin-orbit coupling with the Sr-87
-optical lattice clock, the primary form of decoherence is through
-decay of individual spins, i.e.~with jump operators $\sigma_-^{(j)}$.
-The two axis twisting Hamiltonians $H_{\t{TAT}}^{\z,\y}$ and
-$H_{\t{TAT}}^{\z,\y}$, however, are actually realized in a rotating
-frame, which implies that we must transform these jump operators
-appropriately.  Specifically, the TAT protocol begins with a
-time-dependent transverse-field Hamiltonian
-\begin{align}
-  H = \chi S_\z^2 + \Omega\p{t} S_\x,
-  &&
-  \Omega\p{t} = \beta\omega \cos\p{\omega t},
-\end{align}
-where $\beta$ is the modulation index of the drive, and the drive
-frequency $\omega\gg N\chi$, with $N$ the total number of spins.
-Moving into the rotating frame of $\Omega\p{t}S_\x$ subtracts this
-term from the Hamiltonian, and transforms operators $\O$ as
-\begin{align}
-  \O \to U\p{t}^\dag \O U\p{t},
-  &&
-  U\p{t} \equiv \exp\sp{-i\int_0^t\d\tau~\Omega\p{\tau}S_\x}
-  = \exp\sp{-i\beta\sin\p{\omega t}S_\x},
-\end{align}
-In particular, the operators $\tilde S_\pm\equiv-S_\z\pm iS_\y$
-transform simply as
-\begin{align}
-  \tilde S_\pm
-  \to U^\dag \tilde S_\pm U
-  = e^{\pm i\beta\sin\p{\omega t}} \tilde S_\pm.
-\end{align}
-For any operator $\O$ and drive frequency $\omega\gg\norm{\O}$, where
-$\norm{\O}$ denotes the operator norm of $\O$, we can generally make a
-secular approximation to say
-\begin{align}
-  e^{\pm im\beta\sin\p{\omega t}} \O
-  = \sum_{n=-\infty}^\infty J_n\p{\pm m\beta} e^{in\omega t} \O
-  \approx J_0\p{\pm m\beta} \O
-  = J_0\p{m\beta} \O,
-\end{align}
-where $J_n$ is the $n$-th order Bessel function of the first kind.
-Expanding $S_\z^2=\f14\p{\tilde S_+ + \tilde S_-}^2$, one can thus
-work out that the effective Hamiltonian in the rotating frame of the
-drive is
-\begin{align}
-  H_{\t{eff}}
-  \approx \f{\chi}{2} \p{\sp{1+J_0\p{2\beta}} S_\z^2
-    + \sp{1-J_0\p{2\beta}} S_\y^2}.
-\end{align}
-Driving with a modulation index $\beta$ for which
-$J_0\p{2\beta}=\pm1/3$ then gives us the effective two-axis twisting
-Hamiltonians
-\begin{align}
-  H_{\t{eff}}^{(+)}
-  = \f{\chi}{3} \p{2 S_\z^2 + S_\y^2}
-  \simeq \f{\chi}{3} \p{S_\z^2 - S_\x^2},
-  &&
-  H_{\t{eff}}^{(-)}
-  = \f{\chi}{3} \p{S_\z^2 + 2 S_\y^2}
-  \simeq \f{\chi}{3} \p{S_y^2 - S_\x^2},
-\end{align}
-where $\simeq$ denotes equality up to overall factors of
-$\v S^2=S_\z^2+S_\x^2+S_\y^2$ which are irrelevant in the absence of
-coherent coupling between states with different net spin.  In a
-similar spirit, one can work out that single-spin operators transverse
-to the $\x$-axis transform as
-\begin{align}
-  \tilde\sigma_\pm^{(j)}
-  \equiv \f12\p{-\sigma_\z^{(j)}\pm i\sigma_\y^{(j)}}
-  \to U^\dag \tilde\sigma_\pm^{(j)} U
-  = e^{\pm i\beta\sin\p{\omega t}} \tilde\sigma_\pm^{(j)}
-  \approx J_0\p{\beta} \tilde\sigma_\pm^{(j)},
-\end{align}
-which implies
-\begin{align}
-  \sigma_\z^{(j)} \to J_0\p{\beta} \sigma_\z^{(j)},
-  &&
-  \sigma_\pm^{(j)}
-  \to \f12\sp{1 \pm J_0\p{\beta}} \sigma_+^{(j)}
-  + \f12\sp{1 \mp J_0\p{\beta}} \sigma_-^{(j)}.
-\end{align}
-The TAT protocol about $\p{\z,\y}$ requires
-\begin{align}
-  J_0\p{2\beta} = 1/3
-  \implies \beta \approx 0.906
-  \implies J_0\p{\beta} \approx 0.805,
-\end{align}
-while the TAT protocol about $\p{\y,\x}$ requires
-\begin{align}
-  J_0\p{2\beta} = -1/3
-  \implies
-  \begin{cases}
-    \beta \approx 1.63
-    \implies J_0\p{\beta} \approx 0.440, \\
-    \beta \approx 2.22
-    \implies J_0\p{\beta} \approx 0.099.
-  \end{cases}
-\end{align}
+  \sum_{\v n} T_{\v m\v n;k} \bk{\S_{\v n}}_{t=0}
+  \to \sum_{k=0}^M \f{t^k}{k!}
+  \sum_{\v n} T_{\v m\v n;k} \bk{\S_{\v n}}_{t=0}.
+\end{align}
+The nonzero entries of $T_{\v m\v n;k}$ for $k=0,1,\cdots,M$ now tell
+us which operators we need to keep track of in order to compute the
+expectation value $\bk{\S_{\v m}}$.  Furthermore, after computing all
+relevant matrix elements $T_{\v m\v n;k}$ and expectation values
+$\bk{\S_{\v n}}_{t=0}$, computing $\bk{\S_{\v m}}$ takes only a short
+time independent of system size.  Finally, the relation
+$\S_{\v m}^\dag=\S_{\v m^*}$ for $\v m^*\equiv\p{m_-,m_\z,m_+}$ can be
+exploited to cut both the number of matrix elements $T_{\v m\v n}$ and
+expectation values $\bk{\S_{\v m}}_{t=0}$ that we need to compute
+explicitly roughly in half.
 
@@
 $\sp{S_\mu^m,\sigma_\bmu^{(j)}}_-$, we would like to enforce an
-ordering on products of spin operators, which will allow us to only
+ordering on products of spin operators, which will ensure that we only
 keep track of operators that are linearly independent.  We choose (for
@@
 $E_{mn;\mu}^{(p)}$ in \eqref{eq:E_mn_p} involves a sum over $p$
-mutually dependent intermediate variables!  We therefore devote the
-rest of this section to simplifying our result for the commutator
+mutually dependent intermediate variables, each term of which
+additionally contains a product of $p$ factors!  We therefore devote
+the rest of this section to simplifying our result for the commutator
 $\sp{S_\mu^m,S_\bmu^n}_-$.
@@
 \end{align}
-where $\Gamma$ is a collective spin operator with weight one i.e.~a
+where $\Gamma$ is a collective spin operator with weight one, i.e.~a
 linear combination of $S_\z$, $S_\mu$, and $S_\bmu$.
@@
 
-\section{Changing operator bases}
-
-We wish to change our basis for all exchange-symmetric operators on a
-system of $N$ spins from $S_+^\ell S_\z^m S_-^n$ to
-\begin{align}
-  \S_{\v m} \equiv \sum_{\p{\v j;\v m}} P_{\v j},
+\section{Changing bases for collective-spin operators}
+
+For a collective-spin system composed of many small spins, the
+operators $\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$ keep track
+of $\abs{\v m}$-body operators (for $\abs{\v m}\equiv\sum_\mu m_\mu$),
+but also carry lots of ``baggage'' in the form of $k$-body operators
+with $k<\abs{\v m}$.  To cut down on this overhead, we change our
+basis for collective-spin operators from $\S_{\v m}$ to the
+purely-$\abs{\v m}$-body operators
+\begin{align}
+  \tilde\S_{\v m} \equiv \sum_{\p{\v j;\v m}} P_{\v j},
   &&
@@
 \end{align}
-where $\v m \equiv \p{m_\z,m_+,m_-}$ with $m_\mu$ the number of
-$\sigma_\mu$ operators in each term of $\S_{\v m}$; $\v j$ is an
-ordered list of indices for the spins addressed by $P_{\v j}$; and the
-sum over $\p{\v j;\v m}$ denotes a sum over all possible $\v j$ for
-which all indices in $\v j$ are distinct and
+where $\v m \equiv \p{m_\z,m_+,m_-}\in\mathbb{N}_0^3$ with $m_\mu$ the
+number of $\sigma_\mu$ operators in each term of $\S_{\v m}$; $\v j$
+is an ordered list of indices for the spins addressed by $P_{\v j}$;
+and the sum over $\p{\v j;\v m}$ denotes a sum over all possible
+$\v j$ for which all indices in $\v j$ are distinct and
 $\abs{\set{j^\mu_a}} = m_\mu$, such that $j^\mu_a$ indexes the $a$-th
@@
 
-We now expand a product of basis operators $\S_{\v m}$ and $\S_{\v n}$
-into terms that have exactly $s$ spins addressed by both operators:
-\begin{align}
-  \S_{\v m} \S_{\v n}
+We now expand a product of basis operators $\tilde\S_{\v m}$ and
+$\tilde\S_{\v n}$ into terms that have exactly $s$ spins addressed by
+both operators:
+\begin{align}
+  \tilde\S_{\v m} \tilde\S_{\v n}
   = \sum_{s\ge0} \sum_{\p{\v j,\v k;\v m,\v n,s}} P_{\v j} P_{\v k},
@@
 \begin{align}
-  \S_{\v m}\S_{\v n}
+  \tilde\S_{\v m} \tilde\S_{\v n}
   = \sum_{s\ge0} \sum_{\p{\v r;\v m,\v n,s}} f_{\v m\v n\v r}
@@
 \begin{align}
-  \S_{\v m} \S_{\v n}
+  \tilde\S_{\v m} \tilde\S_{\v n}
   = \sum_{s\ge0} \sum_{\p{\v r;\v m,\v n,s}} f_{\v m\v n\v r}
   \sum_{\p{\v\rho;\v r}} g_{\v\rho} \v\eta_{\v\rho}
-  c_{\v\ell_{\v m\v n\v r\v\rho}\v\rho} \S_{\v\ell_{\v m\v n\v r\v\rho}},
+  c_{\v\ell_{\v m\v n\v r\v\rho}\v\rho}
+  \tilde\S_{\v\ell_{\v m\v n\v r\v\rho}},
   \label{eq:SS_simp}
@@
 \end{align}
+the combinatorial factor
+\begin{align}
+  c_{\v\ell\v\rho}
+  \equiv \p{ N - \sum_\kappa\ell_\kappa }_{\sum_{\mu,\nu}\rho^{\mu\nu}_0}
+\end{align}
+accounts for the number of terms which are equivalent up to a
+permutation of indices involving the identity operators in
+\eqref{eq:Q_K_rho_simp}; and
 the number of $\sigma_\mu$ operators in each term of
-$\S_{\v\ell_{\v m\v n\v r\v\rho}}$ is
+$\tilde\S_{\v\ell_{\v m\v n\v r\v\rho}}$ is
 \begin{align}
@@
   = m_\mu + n_\mu - \sum_\kappa \p{r_{\mu\kappa}+r_{\kappa\mu}}
-  + \sum_{\alpha,\beta} \rho^{\alpha\beta}_\mu;
-\end{align}
-and the combinatorial factor
-\begin{align}
-  c_{\v\ell\v\rho}
-  \equiv \p{ N - \sum_\kappa\ell_\kappa }_{\sum_{\mu,\nu}\rho^{\mu\nu}_0}
-\end{align}
-accounts for the number of terms which are equivalent up to a
-permutation of indices involving the identity operators in
-\eqref{eq:Q_K_rho_simp}.
+  + \sum_{\alpha,\beta} \rho^{\alpha\beta}_\mu.
+\end{align}
 
@@
 can, however, impose additional restrictions on these sums in such a
-way as to throw out terms which vanish, and keep those which do not.
-Remembering that $s$ counts the number of spins which are addressed by
-both operators $P_{\v j}$ and $P_{\v k}$ in \eqref{eq:SS_PP}, the
-first restriction we can impose comes from recognizing that for given
-$\S_{\v m},\S_{\v n}$, the overlap $s$ is bounded as
+way as to throw out terms which vanish, and keep only those which do
+not.  Remembering that $s$ counts the number of spins which are
+addressed by both operators $P_{\v j}$ and $P_{\v k}$ in
+\eqref{eq:SS_PP}, the first restriction we can impose comes from
+recognizing that for given $\tilde\S_{\v m},\tilde\S_{\v n}$, the
+overlap $s$ is bounded as
 \begin{align}
@@
 vanish.  Our second restriction on the sums in \eqref{eq:SS_simp} is
-therefore to fix $r_{\mu\nu}=r_{\nu\mu}=0$ for all $\mu,\nu$ with
+therefore to fix $r_{\mu\nu}=0$ for all $\mu,\nu$ with
 $\sigma_\mu\sigma_\nu=0$.  Finally, we can avoid the nullifying effect

### 2018-10-13 multi_body_interactions/papers/theory/paper.tex
minor changes to multi-body text before arXiv upload

@@
   optical lattice, we develop a low-energy effective theory of
-  fermionic AEAs which exhibits emergent SU($N$)-symmetric multi-body
+  fermionic AEAs which exhibits emergent multi-body SU($N$)-symmetric
   interactions, where $N$ is the number of atomic nuclear spin levels.
@@
 \cite{derevianko2011colloquium, katori2011optical, ludlow2015optical}.
-The atoms have a unique, ultra-narrow optical transition between
+These atoms have a unique, ultra-narrow optical transition between
 metastable ${}^1S_0$ and ${}^3P_0$ electronic orbital states, i.e.~the
@@
 these atoms' degrees of freedom, including the electronic orbital,
-nuclear spin, and motional states, can be fully controlled in a 3-D
-lattice with high fidelity \cite{daley2008quantum,
+nuclear spin, and motional states, can be fully controlled with high
+fidelity in a 3-D lattice \cite{daley2008quantum,
   gorshkov2009alkalineearthmetal, daley2011quantum,
@@
 extraordinary potential to provide novel insights by circumventing
-severe sign problems which plague classical simulations of strongly
-interacting fermionic systems \cite{wieseu.-j.2013ultracold,
+e.g.~severe sign problems which plague classical simulations of
+strongly interacting fermionic systems \cite{wieseu.-j.2013ultracold,
   wu2003exact}.
@@
 neutral bosons prepared in a single hyperfine state, our work deals
-for the first time with fermionic atoms with internal degrees of
-freedom and multiple collisional parameters.  Nonetheless, due to the
-SU($N$) symmetry of these collisions, we are able to find a simple way
-to express effective multi-body Hamiltonians and fully characterize
-their low-lying many-body energy eigenstates.  While some past work
-has detected experimental signatures of multi-body interactions in the
+for the first time with fermions that have internal degrees of freedom
+and multiple collisional parameters.  Nonetheless, due to the SU($N$)
+symmetry of these collisions, we are able to find a simple way to
+express effective multi-body Hamiltonians and fully characterize their
+low-lying many-body energy eigenstates.  While some past work has
+detected experimental signatures of multi-body interactions in the
 form of quantum phase revivals \cite{will2010timeresolved}, we perform
@@
 density-dependent atomic energy spectra performed in
-ref.~\cite{goban2018emergence}, similarly to the measurements of
+ref.~\cite{goban2018emergence}, similarly to the measurements with
 bosons performed in ref.~\cite{mark2011precision}.
@@
 main result of effective multi-body interactions between these atoms.
-In section \ref{sec:low_energy}, we discuss our method for deriving a
+In section \ref{sec:low_energy} we discuss our method for deriving a
 low-energy effective theory, provide a perturbative expansion for the
@@
 (i.e.~projected onto a quantization axis) and orbital state
-$s\in\set{\g,\e}$; and $E_n$ is the energy of a single atom on site
-$i$ in motional state $n$.  In a harmonic trap approximation we would
-have $E_n=\p{3/2+n_x+n_y+n_z}\omega$ for an on-site angular trap
-frequency $\omega$ (with the reduced Planck constant $\hbar=1$
-throughout this paper), but in general the aharmonicity of the lattice
-potential will cause a non-negligible shift in motional state
-energies.
+$s\in\set{\g,\e}$; and $E_n$ is the energy of a single atom in
+motional state $n$.  In a harmonic trap approximation we would have
+$E_n=\p{3/2+n_x+n_y+n_z}\omega$ for an on-site angular trap frequency
+$\omega$ (with the reduced Planck constant $\hbar=1$ throughout this
+paper), but in general the aharmonicity of the lattice potential will
+cause a non-negligible shift in motional state energies.
 
@@
 \end{align}
-where $m_A$ is the mass of a single atom.  Defining the coupling
-tensor
+where $m_A$ is the mass of a single atom.  Defining for brevity
 \begin{align}
@@
 \end{align}
-For nuclear spins $\mu,\nu$, the coupling tensor $G^{qr}_{st}$ gives
-the coupling constant between the two-atom states
+For nuclear spins $\mu,\nu$, the symbol $G^{qr}_{st}$ gives the
+coupling constant between the two-atom states
 $(\mu,q)+(\nu,r)\leftrightarrow(\mu,s)+(\nu,t)$.
@@
 where for brevity we will henceforth suppress the identical site index
-($i$) on all operators, and implicitly sum over all indices in a
-summand which do not have a fixed value.  We may also at times
-suppress motional state indices on the overlap integral
+($i$) on all operators, and implicitly sum over all free indices in a
+summand (i.e.~indices which do not have a fixed value).  We may also
+at times suppress motional state indices on the overlap integral
 $K^{k\ell}_{mn}$, in which case the suppressed indices are implicitly
@@
 two-body overlap integrals are real, such that
-$K^{k\ell}_{mn}={K^{k\ell}_{mn}}^*=K^{mn}_{k\ell}$.  This choice is
-always possible due to the analytic properties of Wannier orbitals
-$\phi_{in}$ \cite{kohn1959analytic}.
+$K^{k\ell}_{mn}={K^{k\ell}_{mn}}^*=K^{mn}_{k\ell}$.  The existence of
+such a gauge is guaranteed by the analytic properties of the Wannier
+orbitals $\phi_{in}$ \cite{kohn1959analytic}.
 
@@
 excited single-particle motional states fails to reproduce the
-observed spectra of multiply-occupied lattice sites
-\cite{goban2018emergence}.  The reason for this failure lies in the
-fact that the bare two-body Hamiltonian $H_{\t{int}}$ in
-\eqref{eq:H_int} is not diagonal in the basis of single-particle
-motional states.  Even at zero temperature, therefore, the presence of
-excited motional states modifies many-body atomic energy spectra when
-off-diagonal elements of $H_{\t{int}}$ approach the scale of low-lying
-single-particle motional state energy gaps.  To address this problem,
-we develop a low-energy effective theory for ultracold two-level
-fermionic atoms on a lattice.  As our work is motivated by progress in
-experiments which initialize all atoms in identical (i.e.~ground)
-orbital and motional states, to simplify our theory we assume that any
-$N$ atoms on a single lattice site occupy $N$ nuclear spin states.
-Multiple occupation of a single nuclear spin state on the same lattice
-site is initially forbidden by fermionic statistics, and subsequently
-cannot be violated in the absence of tunneling and hyperfine coupling.
+observed spectra of multiply-occupied lattice sites.  The reason for
+this failure lies in the fact that the bare two-body Hamiltonian
+$H_{\t{int}}$ in \eqref{eq:H_int} is not diagonal in the basis of
+single-particle motional states.  Even at zero temperature, therefore,
+the presence of excited motional states modifies many-body atomic
+energy spectra when the magnitude of off-diagonal elements of
+$H_{\t{int}}$ approach the scale of low-lying single-particle motional
+state energy gaps.  To address this problem, we develop a low-energy
+effective theory for ultracold two-level fermionic atoms on a lattice.
+As our work is motivated by progress in experiments which initialize
+all atoms in identical (i.e.~ground) orbital and motional states, to
+simplify our theory we assume that any $N$ atoms on a single lattice
+site occupy $N$ nuclear spin states.  Multiple occupation of a single
+nuclear spin state on the same lattice site is initially forbidden by
+fermionic statistics, and subsequently cannot be violated in the
+absence of tunneling and hyperfine coupling.
 
@@
 where $H_M$ is an $M$-body Hamiltonian, $I$ is the total nuclear spin
-of each atom (e.g.~$I=9/2$ for ${}^{87}$Sr), and the sum ends at
+of each atom (e.g.~$I=9/2$ for ${}^{87}$Sr), and the sum terminates at
 $2I+1$ because this is the largest number of atoms which may initially
 occupy a single lattice site.  Focusing on the low-lying orbital state
-excitations of this theory, we find that we can express multi-body
-Hamiltonians in the simple form
+excitations of this theory, we find that SU($N$) symmetry allows us to
+express multi-body Hamiltonians in the simple form
 \begin{align}
@@
 and the sum is performed over all choices of nuclear spins $\mu_j$
-with $j=1,\cdots,M$ for which all $\mu_j$ are distinct, or
+with $j=1,2,\cdots,M$ for which all $\mu_j$ are distinct, or
 equivalently all choices of $\mu_j$ for which the set $\set{\mu_j}$
@@
 the many-body problem more tractable by greatly reducing the
-dimensionality of our target Hilbert space.
+dimensionality of our the relevant Hilbert space.
 
@@
 \end{align}
-where $\sp{X,Y}_+\equiv XY+YX$.  The naive idea of applying
-$H_{\t{int}}$ directly on the single-particle motional ground states
-and neglecting all excited states is thus equivalent to truncating our
-expansion for $H_{\t{int}}^{\t{eff}}$ at first order.  In addition to
-this first order term, the expansion involves {\it effective
-  corrections} to the action of $H_{\t{int}}$ on the single-particle
-motional ground states (i.e.~on $\H_{\t{ground}}^{\t{single}}$) in the
-form of higher-order terms with {\it intermediate} or {\it virtual
-  occupation} of excited states, manifest in $\I$.
+where $\sp{X,Y}_+\equiv XY+YX$.  The naive idea of neglecting all
+excited states and using $H_{\t{int}}$ directly to describe the
+spectrum of single-particle motional ground states is thus equivalent
+to truncating our expansion for $H_{\t{int}}^{\t{eff}}$ at first
+order.  In addition to this first order term, the expansion involves
+{\it effective corrections} to the action of $H_{\t{int}}$ on the
+single-particle motional ground states (i.e.~on
+$\H_{\t{ground}}^{\t{single}}$) in the form of higher-order terms with
+{\it intermediate} or {\it virtual occupation} of excited states,
+manifest in $\I$.
 
@@
 operator equations, and that the sums over virtual states are
-performed over a basis for the orthogonal complement of
+performed over a basis for the orthogonal complement of the subspace
 $\H_{\t{ground}}^{\t{single}}$, rather than a basis for the orthogonal
@@
     written next to each vertex; we will generally not include these
-    features, as they are not necessary to uniquely identify a
-    diagram.  We will also drop explicit appearances of the
-    ground-state projector $\P_0$ in our expressions, with the
+    features, as they are not necessary to uniquely identify the term
+    represented by a diagram.  We will also drop explicit appearances
+    of the ground-state projector $\P_0$ in our expressions, with the
     understanding that the low-energy effective theory implicitly
@@
 where $\alpha_3^{(2)} \equiv \sum_{n>0} K_n^2/E_n$, and the preceding
-negative sign is as prescribed by $H_{\t{int}}^{(2)}$ in
+minus sign is as prescribed by $H_{\t{int}}^{(2)}$ in
 \eqref{eq:H_int_1_2}.
@@
 compactly enumerate and evaluate all three-body diagrams at third
-order, we introduce an expanded coupling tensor
+order, we introduce an expanded coupling symbol
 \begin{align}
@@
 At the cost of introducing an additional sum over new nuclear spin
-indices, the expanded coupling tensor allows us to collect together
-diagrams which have identical graph topologies, but address different
+indices, the expanded coupling symbol allows us to collect together
+diagrams which have the same graph topology, but represent different
 matrix elements of the effective Hamiltonian due to the exchange of
@@
 $H_{\t{int}}^{(3)}$ in \eqref{eq:H_int_3}, these diagrams have an
-associated negative sign if they contain only one excited virtual
-state, and a factor of $1/2$ if they contain a virtual ground state.
+associated minus sign if they contain only one excited virtual state,
+and a factor of $1/2$ if they contain a virtual ground state.
 Remembering that counter-terms are $\O\p{G^2}$, there are additionally
@@
 in \eqref{eq:H_3_3_O} and \eqref{eq:H_3_3_counter_term}, which results
-in a contribution to the effective Hamiltonian which is equal to
+in a contribution to the effective Hamiltonian that is equal to
 \eqref{eq:H_3_3_cancellation}.  The contribution from the
@@
 As prescribed by $H_{\t{int}}^{(3)}$ in \eqref{eq:H_int_3}, however,
-the second and third of these diagrams pick up a factor of $-1/2$, and
-the sum of the disconnected diagrams vanishes.
+the second and third of these diagrams pick up a factor of $-1/2$, so
+the sum over disconnected diagrams vanishes.
 
@@
 $N\ge M$ atoms occupying nuclear spins $\N=\set{\mu_j}$ for
-$j=1,\cdots,N$, the $M$-body Hamiltonian $H_M$ has a single ground
+$j=1,2,\cdots,N$, the $M$-body Hamiltonian $H_M$ has a single ground
 state $\ket{\N,0}$, and a singly-excited state $\ket{\N,+}$ which is
@@
     Hamiltonians $H_M$.  Many-body energy eigenstates are labeled by
-    the nuclear spins they occupy (i.e.~$\N$ with $\abs{\N}=N$) and
-    whether they are in the orbital ground ($0$), singly-excited
+    the nuclear spins they occupy (i.e.~$\N$ with $N\equiv\abs{\N}$)
+    and whether they are in an orbital ground ($0$), singly-excited
     symmetric ($+$), or singly-excited asymmetric ($-$) state.  The
@@
 is proportional to the Clebsch-Gordan coefficient
-$\bk{I,\mu;1,0|I,\mu}\sim\mu$ for a photon-induced
+$\bk{I,\mu;1,0|I,\mu}\propto\mu$ for a photon-induced
 nuclear-spin-conserving orbital state transition of an atom with
-nuclear spin $\mu$.  We therefore define the bare Rabi frequency
+nuclear spin $\mu$.  We therefore define the ``bare'' Rabi frequency
 $\Omega_0\equiv\Omega_\mu/\mu$ to explicitly factor out dependence on
@@
 $\ket{\N,0}$ with nuclear spins $\N\equiv\set{\mu_j}$ for
-$j=1,\cdots,N$.  If we red-detune the interrogation laser by $\delta$
-from a many-body orbital state excitation energy, i.e.~set
+$j=1,2,\cdots,N$.  If we red-detune the interrogation laser by
+$\delta$ from a many-body orbital state excitation energy, i.e.~set
 $\Delta=\Delta_{NX}-\delta$ for $\Delta_{NX}\equiv E_{NX}-E_{N,0}$ and

### 2018-10-22 soc_squeezing/scripts/correlator_methods.py
changed convention for decoherence vector ordering to PZM changed convention (involving factors of 2) for dephasing rate for consistency with literature

@@ -303,7 +303,7 @@ def convert_zxy(vec_zxy, mu):
 # diagonal terms of single-spin decoherence
 def op_image_decoherence_diag_individual(op, SS, dec_vec, mu):
     ll, mm, nn = op
-    D_z, D_p, D_m = abs(np.array(dec_vec))**2
+    D_p, D_z, D_m = abs(np.array(dec_vec))**2
     if mu == 1:
         D_mu, D_nu = D_p, D_m
     else:
@@ -336,7 +336,7 @@ def op_image_decoherence_diag_individual(op, SS, dec_vec, mu):
 # single-spin decoherence "Q" cross term
 def op_image_decoherence_Q_individual(op, SS, dec_vec, mu):
     ll, mm, nn = op
-    g_z, g_p, g_m = dec_vec
+    g_p, g_z, g_m = dec_vec
     if mu == 1:
         g_mu, g_nu = g_p, g_m
     else:
@@ -381,7 +381,7 @@ def op_image_decoherence_Q_individual(op, SS, dec_vec, mu):
 # diagonal terms of collective-spin decoherence
 def op_image_decoherence_diag_collective(op, SS, dec_vec, mu):
     ll, mm, nn = op
-    D_z, D_p, D_m = abs(np.array(dec_vec))**2
+    D_p, D_z, D_m = abs(np.array(dec_vec))**2
     if mu == 1:
         D_mu, D_nu = D_p, D_m
     else:
@@ -416,7 +416,7 @@ def op_image_decoherence_diag_collective(op, SS, dec_vec, mu):
 # collective-spin decoherence "Q" cross term
 def op_image_decoherence_Q_collective(op, SS, dec_vec, mu):
     ll, mm, nn = op
-    g_z, g_p, g_m = dec_vec
+    g_p, g_z, g_m = dec_vec
     if mu == 1:
         g_mu, g_nu = g_p, g_m
     else:
@@ -469,12 +469,12 @@ def op_image_decoherence_Q_collective(op, SS, dec_vec, mu):
 
 # convert decoherence rates and transformation matrix to decoherence vectors
 def get_dec_vecs(dec_rates, dec_mat):
+    if dec_rates == []: return []
     dec_vecs = []
     for jj in range(3):
         dec_vec_g = dec_mat[:,jj] * np.sqrt(dec_rates[0][jj])
         dec_vec_G = dec_mat[:,jj] * np.sqrt(dec_rates[1][jj])
[…]

### 2018-10-22 multi_body_interactions/papers/theory/paper.tex
added keyword, updated acknowledgements statement, and fixed typo in multi-body theory paper

@@
 
-\keywords{\it ultracold atoms, optical lattices, SU(N) magnetism}
+\keywords{\it ultracold atoms, optical lattices, SU(N) magnetism,
+  multi-body interactions}
 
@@
 L.~Radzihovsky; and in particular close correspondence and technical
-contributions from A.~Goban.  This work is supported by NIST, DARPA,
-W911NF-16-1-0576 through ARO, AFOSR, AFOSR-MURI, NSF-1734006, NASA,
-and CTQM.
+contributions from A.~Goban.  This work is supported by the Air Force
+Office of Scientific Research (AFOSR) grant FA9550-18-1-0319; the
+AFOSR Multidisciplinary University Research Initiative (MURI) grant;
+the Defense Advanced Research Projects Agency (DARPA) and Army
+Research Office (ARO) grant W911NF-16-1-0576; the National Science
+Foundation (NSF) grant PHY-1820885; JILA-NSF grant PFC-173400; and the
+National Institute of Standards and Technology (NIST).
 
@@
   H_2' \equiv \f12 K' \sum {G'}^{qr}_{st}
-  \leq\c_{\mu s}^\dag \c_{\nu t}^\dag \c_{\nu r} \c_{\mu q},
+  \c_{\mu s}^\dag \c_{\nu t}^\dag \c_{\nu r} \c_{\mu q},
   \label{eq:H_2_prime}

### 2018-11-04 soc_squeezing/cumulant_expansions.tex
changed convention for decoherence vector ordering to PZM changed convention (involving factors of 2) for dephasing rate for consistency with literature
added todo line into the cumulant expansions text
change section on the general commutator of spin operators to a section on the general product

@@
 $\uv n$ in the plane orthogonal to the mean spin vector $\bk{\v S}$.
-This squeezing parameter is entirely determined by components
-$\bk{S_\alpha}$ of the mean spin vector in addition to collective
-spin-spin correlators of the form $\bk{S_\alpha S_\beta}$; our task is
-therefore to compute these expectation values in all scenarios of
-interest.
+This squeezing parameter is entirely determined by one- and two-spin
+correlators of the form $\bk{S_\alpha}$ and $\bk{S_\alpha S_\beta}$;
+our task is therefore to compute these expectation values in all
+scenarios of interest.
 
@@
 matrix elements $T_{\v m\v n}$ of the time derivative operator $d/dt$
-are worked out Appendices \ref{sec:comm_general},
+are worked out Appendices \ref{sec:prod_general},
 \ref{sec:decoherence_single}, and \ref{sec:decoherence_collective}.
@@
 \label{sec:identities}
+% TODO: change { \sigma_\mu } --> { s_\mu }
 
@@
 
-\section{Commutation relations between ordered products of collective
-  spin operators}
-\label{sec:comm_general}
-
-We will generally be concerned with ordered products of collective
-spin operators of the form $S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$ for
-$\bmu\equiv-\mu\in\set{+1,-1}$.  The most general (anti-)commutator we
-need to consider is then
-\begin{align}
-  \sp{S_\mu^p \p{\mu S_\z}^q S_\bmu^r,
-    S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_\pm
-  = \S^{pqr}_{\ell mn;\mu} \pm \S^{\ell mn}_{pqr;\mu},
-\end{align}
-where
+\section{Product of arbitrary ordered collective spin operators}
+\label{sec:prod_general}
+
+The most general product of collective spin operators that we need to
+compute is
 \begin{align}
@@
 \end{align}
-for
+where
 \begin{align}
@@
 \end{align}
-defined in terms of $Z_{r\ell;\mu}^{(k)}$ and $\zeta_{r\ell}^{ka}$ as
-given in \eqref{eq:Z_mnp} and \eqref{eq:zeta_mnpq}.
+is defined in terms of $Z_{r\ell;\mu}^{(k)}$ and $\zeta_{r\ell}^{ka}$
+as respectively given in \eqref{eq:Z_mnp} and \eqref{eq:zeta_mnpq}.
+The (anti-)commutator of two ordered products of collective spin
+operators is then
+\begin{align}
+  \sp{S_\mu^p \p{\mu S_\z}^q S_\bmu^r,
+    S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_\pm
+  = \S^{pqr}_{\ell mn;\mu} \pm \S^{\ell mn}_{pqr;\mu}.
+\end{align}
 
@@
 \begin{align}
-  \D\p{\sigma_\z/\sqrt2; \O}
-  = \f12 \sum_j\p{\sigma_\z^{(j)} \O \sigma_\z^{(j)}
+  \D\p{\sigma_\z; \O}
+  = \sum_j\p{\sigma_\z^{(j)} \O \sigma_\z^{(j)}
     - \f12\sp{\sigma_\z^{(j)} \sigma_\z^{(j)},\O}_+}
-  = \f12 \sum_j \sigma_\z^{(j)} \O \sigma_\z^{(j)} - S \O,
-\end{align}
-where the factor $\sqrt2$ is inserted as
-e.g.~$\D\p{\sigma_\z/\sqrt2;S_\mu}=-S_\mu$, such that the above
-expression generates decay of single-weight transverse collective spin
-operators at a unit rate.  From \eqref{eq:san_z_z}, we then have
-\begin{align}
-  \D\p{\sigma_\z/\sqrt2; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
-  = -\p{\ell+n} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n
-  + 2 \ell n S_\mu^{\ell-1} \p{S + \mu S_\z}
+  = \sum_j \sigma_\z^{(j)} \O \sigma_\z^{(j)} - 2 S \O.
+\end{align}
+From \eqref{eq:san_z_z}, we then have
+\begin{align}
+  \D\p{\sigma_\z; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
+  = -2\p{\ell+n} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n
+  + 4 \ell n S_\mu^{\ell-1} \p{S + \mu S_\z}
   \p{-1 + \mu S_\z}^m S_\bmu^{n-1}.
@@
   &\qquad + \mu \Gamma_{\z,\mu}^{(-)}
-  S_\mu^{\ell+1} \p{\mu S_\z}
+  S_\mu^{\ell+1} \mu S_\z
   \sp{\p{1+\mu S_\z}^m - \p{\mu S_\z}^m} S_\bmu^n,
@@
   \sigma_\kappa^{\otimes\sum_{\mu,\nu}\tilde\rho^{\mu\nu}_\kappa}
-  = \sum_{\tilde{\v\rho}} g_{\tilde{\v\rho}} \v\eta_{\tilde{\v\rho}}
+  \simeq \sum_{\tilde{\v\rho}} g_{\tilde{\v\rho}} \v\eta_{\tilde{\v\rho}}
   \1^{\otimes\sum_{\mu,\nu}\tilde\rho^{\mu\nu}_0}

### 2018-11-08 soc_squeezing/cumulant_expansions.tex
changed all text and testing scripts to use spin-1/2 operators instead of pauli operators

@@
   &&
-  \Gamma \equiv \gamma_\z + \lambda,
+  \Gamma \equiv \f12\gamma_\z + \lambda,
   &&
@@
 \label{sec:identities}
-% TODO: change { \sigma_\mu } --> { s_\mu }
 
@@
 operator identities; we collect and derive some basic identities here
-for reference.  The elementary commutation relations between spin
+for reference.  Defining the individual spin-1/2 operators
+\begin{align}
+  s_{n=\x,\y,\z} \equiv \f12 \sigma_n,
+  &&
+  s_\pm \equiv s_\x \pm i s_\y,
+  \label{eq:single_spin}
+\end{align}
+in terms of Pauli operators $\sigma_n$, we denote the operator which
+acts with $s_n$ on a spin indexed by $j$ and trivially (i.e.~with the
+identity $\1$) on all other spins by $s_n^{(j)}$.  We then define the
+collective spin operators
+\begin{align}
+  S_n \equiv \sum_j s_n^{(j)}.
+  \label{eq:collective_spin}
+\end{align}
+Despite our working definition in \eqref{eq:collective_spin}, the
+identities we will derive for collective spin operators alone apply
+just as well to large-spin operators that cannot be expressed as the
+sum of individual spin-1/2 operators.
+
+The elementary commutation relations between spin
 operators are, with $\bmu\equiv-\mu\in\set{+1,-1}$ for brevity,
 \begin{align}
-  \f12 \sp{\sigma_\z^{(j)},\sigma_\mu^{(k)}}_-
-  &= \delta_{jk} \mu\sigma_\mu^{(j)},
+  \sp{s_\z^{(j)},s_\mu^{(k)}}_-
+  &= \delta_{jk} \mu s_\mu^{(j)},
   &
-  \sp{S_\z,\sigma_\mu^{(j)}}_-
-  &= \f12 \sp{\sigma_\z^{(j)},S_\mu}_- = \mu\sigma_\mu^{(j)},
+  \sp{S_\z,s_\mu^{(j)}}_-
+  &= \sp{s_\z^{(j)},S_\mu}_- = \mu s_\mu^{(j)},
   &
@@
   \label{eq:comm_z_base} \\
-  \sp{\sigma_\mu^{(j)},\sigma_\bmu^{(k)}}_-
-  &= \delta_{jk} \mu \sigma_\z^{(j)},
+  \sp{s_\mu^{(j)},s_\bmu^{(k)}}_-
+  &= \delta_{jk} 2 \mu s_\z^{(j)},
   &
-  \sp{S_\mu,\sigma_\bmu^{(j)}}_-
-  &= \sp{\sigma_\mu^{(j)},S_\bmu}_- = \mu \sigma_\z^{(j)},
+  \sp{S_\mu,s_\bmu^{(j)}}_-
+  &= \sp{s_\mu^{(j)},S_\bmu}_- = 2 \mu s_\z^{(j)},
   &
@@
 \begin{align}
-  \p{\mu S_\z}^m \sigma_\mu^{(j)}
-  = \p{\mu S_\z}^{m-1} \sigma_\mu^{(j)} \p{1 + \mu S_\z}
-  = \p{\mu S_\z}^{m-2} \sigma_\mu^{(j)} \p{1 + \mu S_\z}^2
+  \p{\mu S_\z}^m s_\mu^{(j)}
+  = \p{\mu S_\z}^{m-1} s_\mu^{(j)} \p{1 + \mu S_\z}
+  = \p{\mu S_\z}^{m-2} s_\mu^{(j)} \p{1 + \mu S_\z}^2
   = \cdots
-  = \sigma_\mu^{(j)} \p{1 + \mu S_\z}^m,
+  = s_\mu^{(j)} \p{1 + \mu S_\z}^m,
   \label{eq:push_z_mu_Ss}
@@
 \begin{align}
-  \mu\sigma_\z^{(j)} S_\mu^m
-  = S_\mu \mu\sigma_\z^{(j)} S_\mu^{m-1} + 2\sigma_\mu^{(j)} S_\mu^{m-1}
+  \mu s_\z^{(j)} S_\mu^m
+  = S_\mu \mu s_\z^{(j)} S_\mu^{m-1} + s_\mu^{(j)} S_\mu^{m-1}
   = \cdots
-  = 2m\sigma_\mu^{(j)} S_\mu^{m-1} + S_\mu^m \mu \sigma_\z^{(j)},
+  = ms_\mu^{(j)} S_\mu^{m-1} + S_\mu^m \mu s_\z^{(j)},
   \label{eq:push_z_mu_sS}
 \end{align}
-where we will generally find it nicer to express our results in terms
-of $\mu\sigma_\z^{(j)}$ and $\mu S_\z$ rather than $\sigma_\z^{(j)}$
-and $S_\z$.  Summing over the single-spin index $j$ in both of these
-cases gives us the purely collective-spin versions of these
-identities:
+where we will generally find it nicer to express results in terms of
+$\mu s_\z^{(j)}$ and $\mu S_\z$ rather than $s_\z^{(j)}$ and $S_\z$.
+Summing over the single-spin index $j$ in both of these cases gives us
+the purely collective-spin versions of these identities:
 \begin{align}
@@
 To find commutation relations between powers of transverse collective
-spin operators, we compute (with $\bmu\equiv-\mu\in\set{+1,-1}$)
-\begin{align}
-  S_\mu^m \sigma_\bmu^{(j)}
-  &= S_\mu^{m-1}\sigma_\bmu^{(j)} S_\mu
-  + S_\mu^{m-1} \mu\sigma_\z^{(j)} \\
-  &= S_\mu^{m-2} \sigma_\bmu^{(j)} S_\mu^2
-  + S_\mu^{m-2} \mu\sigma_\z^{(j)} S_\mu
-  + S_\mu^{m-1} \mu\sigma_\z^{(j)} \\
-  &= \sigma_\bmu^{(j)} S_\mu^m
-  + \sum_{k=0}^{m-1} S_\mu^k \mu\sigma_\z^{(j)} S_\mu^{m-k-1}
+spin operators, we first compute
+\begin{align}
+  S_\mu^m s_\bmu^{(j)}
+  &= S_\mu^{m-1}s_\bmu^{(j)} S_\mu
+  + S_\mu^{m-1} 2\mu s_\z^{(j)} \\
+  &= S_\mu^{m-2} s_\bmu^{(j)} S_\mu^2
+  + S_\mu^{m-2} 2\mu s_\z^{(j)} S_\mu
+  + S_\mu^{m-1} 2\mu s_\z^{(j)} \\
+  &= s_\bmu^{(j)} S_\mu^m
+  + \sum_{k=0}^{m-1} S_\mu^k 2\mu s_\z^{(j)} S_\mu^{m-k-1}
   \label{eq:push_mu_Ss_start}.
@@
 While \eqref{eq:push_mu_Ss_start} gives us the commutator
-$\sp{S_\mu^m,\sigma_\bmu^{(j)}}_-$, we would like to enforce an
-ordering on products of spin operators, which will ensure that we only
-keep track of operators that are linearly independent.  We choose (for
-now) to impose an ordering with all $\sigma_\bmu^{(j)}$ operators on
-the left, and all $\sigma_\z^{(j)}$ operators on the right, as such an
-ordering will prove convenient for the calculations in this section.
-This choice of ordering compels us to expand
-\begin{align}
-  \sum_{k=0}^{m-1} S_\mu^k \mu\sigma_\z^{(j)} S_\mu^{m-k-1}
+$\sp{S_\mu^m,s_\bmu^{(j)}}_-$, we would like to enforce an ordering on
+products of spin operators, which will ensure that we only keep track
+of operators that are linearly independent.  We choose (for now) to
+impose an ordering with all $s_\bmu^{(j)}$ operators on the left, and
+all $s_\z^{(j)}$ operators on the right, as such an ordering will
+prove convenient for the calculations in this section.  This choice of
+ordering compels us to expand
+\begin{align}
+  \sum_{k=0}^{m-1} S_\mu^k 2\mu s_\z^{(j)} S_\mu^{m-k-1}
   &= \sum_{k=0}^{m-1} S_\mu^k
-  \sp{2\p{m-k-1} \sigma_\mu^{(j)} S_\mu^{m-k-2}
-    + S_\mu^{m-k-1} \mu\sigma_\z^{(j)}} \\
-  &= m \p{m-1} \sigma_\mu^{(j)} S_\mu^{m-2}
-  + m S_\mu^{m-1} \mu\sigma_\z^{(j)},
+  \sp{2\p{m-k-1} s_\mu^{(j)} S_\mu^{m-k-2}
+    + S_\mu^{m-k-1} 2\mu s_\z^{(j)}} \\
+  &= m \p{m-1} s_\mu^{(j)} S_\mu^{m-2}
+  + m S_\mu^{m-1} 2\mu s_\z^{(j)},
 \end{align}
@@
 \begin{align}
-  S_\mu^m \sigma_\bmu^{(j)}
-  = \sigma_\bmu^{(j)} S_\mu^m + m \p{m-1} \sigma_\mu^{(j)} S_\mu^{m-2}
-  + m S_\mu^{m-1} \mu\sigma_\z^{(j)},
+  S_\mu^m s_\bmu^{(j)}
+  = s_\bmu^{(j)} S_\mu^m + m \p{m-1} s_\mu^{(j)} S_\mu^{m-2}
+  + m S_\mu^{m-1} 2\mu s_\z^{(j)},
   \label{eq:push_mu_Ss}
@@
 order, which means pushing all $S_\z$ operators to the right and, for
-the purposes of this calculation, all $S_\bmu$ operators on the left.
+the purposes of this calculation, all $S_\bmu$ operators to the left.
 We begin by pushing $S_\mu^k$ to the left of $S_\z$, which takes
@@
 mutually dependent intermediate variables, each term of which
-additionally contains a product of $p$ factors!  We therefore devote
+additionally contains a product of $p$ factors.  We therefore devote
 the rest of this section to simplifying our result for the commutator
@@
 sandwiching $S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$.  Up to Hermitian
-conjugation, the unique cases are, for $\mu,\nu\in\set{+1,-1}$ and
-$\bmu\equiv-\mu$,
-\begin{align}
-  \sum_j \sigma_\z^{(j)} \p{\mu S_\z}^m \sigma_\z^{(j)}
-  &= \sum_j \sigma_\z^{(j)} \sigma_\z^{(j)} \p{\mu S_\z}^m
-  = \sum_j \1_j \p{\mu S_\z}^m
-  = 2 S \p{\mu S_\z}^m, \\
-  \sum_j \sigma_\z^{(j)} \p{\mu S_\z}^m \sigma_\nu^{(j)}
-  &= \p{\mu S_\z}^m \sum_j \sigma_\z^{(j)} \sigma_\nu^{(j)}
-  = \p{\mu S_\z}^m \nu S_\nu
-  = \nu S_\nu \p{\mu\nu+\mu S_\z}^m, \\
-  \sum_j \sigma_\nu^{(j)} \p{\mu S_\z}^m \sigma_\nu^{(j)}
-  &= \sum_j \sigma_\nu^{(j)} \sigma_\nu^{(j)} \p{\mu\nu+\mu S_\z}^m
+conjugation, the unique cases are, for $\mu,\nu\in\set{+1,-1}$,
+\begin{align}
+  \sum_j s_\z^{(j)} \p{\mu S_\z}^m s_\z^{(j)}
+  &= \sum_j s_\z^{(j)} s_\z^{(j)} \p{\mu S_\z}^m
+  = \f14 \sum_j \1_j \p{\mu S_\z}^m
+  = \f12 S \p{\mu S_\z}^m, \\
+  \sum_j s_\z^{(j)} \p{\mu S_\z}^m s_\nu^{(j)}
+  &= \p{\mu S_\z}^m \sum_j s_\z^{(j)} s_\nu^{(j)}
+  = \f12 \p{\mu S_\z}^m \nu S_\nu
+  = \f12 \nu S_\nu \p{\mu\nu+\mu S_\z}^m, \\
+  \sum_j s_\nu^{(j)} \p{\mu S_\z}^m s_\nu^{(j)}
+  &= \sum_j s_\nu^{(j)} s_\nu^{(j)} \p{\mu\nu+\mu S_\z}^m
   = 0, \\
-  \sum_j \sigma_\bnu^{(j)} \p{\mu S_\z}^m \sigma_\nu^{(j)}
-  &= \sum_j \sigma_\bnu^{(j)} \sigma_\nu^{(j)} \p{\mu\nu+\mu S_\z}^m
+  \sum_j s_\bnu^{(j)} \p{\mu S_\z}^m s_\nu^{(j)}
+  &= \sum_j s_\bnu^{(j)} s_\nu^{(j)} \p{\mu\nu+\mu S_\z}^m
   = \p{S-\nu S_\z} \p{\mu\nu+\mu S_\z}^m.
@@
 \begin{align}
-  \sum_j \sigma_\z^{(j)} S_\mu^\ell
-  \p{\mu S_\z}^m S_\bmu^n \sigma_\z^{(j)}
-  &= 2\p{S-\ell-n} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n \notag \\
-  &\qquad + 4\ell n S_\mu^{\ell-1} \p{S+\mu S_\z}
+  \sum_j s_\z^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\z^{(j)}
+  &= \f12 \p{S-\ell-n} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n
+  + \ell n S_\mu^{\ell-1} \p{S+\mu S_\z}
   \p{-1+\mu S_\z}^m S_\bmu^{n-1},
   \label{eq:san_z_z} \allowdisplaybreaks \\
-  \sum_j \sigma_\z^{(j)} S_\mu^\ell
-  \p{\mu S_\z}^m S_\bmu^n \sigma_\mu^{(j)}
-  &= \mu S_\mu^{\ell+1} \p{1+\mu S_\z}^m S_\bmu^n
-  + \mu n \p{-2S+2\ell+n-1} S_\mu^\ell
+  \sum_j s_\z^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\mu^{(j)}
+  &= \f12 \mu S_\mu^{\ell+1} \p{1+\mu S_\z}^m S_\bmu^n
+  - \mu n \p{S-\ell-\f12\sp{n-1}} S_\mu^\ell
   \p{\mu S_\z}^m S_\bmu^{n-1} \notag \\
-  &\qquad - 2\mu\ell n\p{n-1} S_\mu^{\ell-1}
+  &\qquad - \mu\ell n\p{n-1} S_\mu^{\ell-1}
   \p{S+\mu S_\z} \p{-1+\mu S_\z}^m S_\bmu^{n-2},
   \label{eq:san_z_mu} \allowdisplaybreaks \\
-  \sum_j \sigma_\z^{(j)} S_\mu^\ell
-  \p{\mu S_\z}^m S_\bmu^n \sigma_\bmu^{(j)}
-  &= -\mu S_\mu^\ell \p{\mu S_\z}^m S_\bmu^{n+1}
-  + 2\mu \ell S_\mu^{\ell-1} \p{S+\mu S_\z} \p{-1+\mu S_\z}^m S_\bmu^n,
+  \sum_j s_\z^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\bmu^{(j)}
+  &= -\f12 \mu S_\mu^\ell \p{\mu S_\z}^m S_\bmu^{n+1}
+  + \mu \ell S_\mu^{\ell-1} \p{S+\mu S_\z} \p{-1+\mu S_\z}^m S_\bmu^n,
   \label{eq:san_z_bmu} \allowdisplaybreaks \\
-  \sum_j \sigma_\mu^{(j)} S_\mu^\ell
-  \p{\mu S_\z}^m S_\bmu^n \sigma_\mu^{(j)}
+  \sum_j s_\mu^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\mu^{(j)}
   &= n S_\mu^{\ell+1} \p{\mu S_\z}^m S_\bmu^{n-1}
@@
   \label{eq:san_mu_mu} \allowdisplaybreaks \\
-  \sum_j \sigma_\mu^{(j)} S_\mu^\ell
-  \p{\mu S_\z}^m S_\bmu^n \sigma_\bmu^{(j)}
+  \sum_j s_\mu^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\bmu^{(j)}
   &= S_\mu^\ell \p{S+\mu S_\z}\p{-1+\mu S_\z}^m S_\bmu^n,
   \label{eq:san_mu_bmu} \allowdisplaybreaks \\
-  \sum_j \sigma_\bmu^{(j)} S_\mu^\ell
-  \p{\mu S_\z}^m S_\bmu^n \sigma_\mu^{(j)}
+  \sum_j s_\bmu^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\mu^{(j)}
   &= S_\mu^\ell \p{S - \ell - n - \mu S_\z}
@@
 The effect of decoherence via uncorrelated decay ($\mu=-1$) or
-excitation ($\mu=1$) of individual spins is described by (with
-$\bmu\equiv-\mu$),
-\begin{align}
-  \D\p{\sigma_\mu; \O}
-  = \sum_j\p{\sigma_\bmu^{(j)} \O \sigma_\mu^{(j)}
-    - \f12\sp{\sigma_\bmu^{(j)} \sigma_\mu^{(j)},\O}_+}
-  = \sum_j \sigma_\bmu^{(j)} \O \sigma_\mu^{(j)}
+excitation ($\mu=1$) of individual spins is described by
+\begin{align}
+  \D\p{s_\mu; \O}
+  = \sum_j\p{s_\bmu^{(j)} \O s_\mu^{(j)}
+    - \f12\sp{s_\bmu^{(j)} s_\mu^{(j)},\O}_+}
+  = \sum_j s_\bmu^{(j)} \O s_\mu^{(j)}
   - S \O + \f{\mu}{2} \sp{S_\z, \O}_+.
@@
 \begin{align}
-  \D\p{\sigma_\bmu; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
+  \D\p{s_\bmu; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
   = S_\mu^\ell \p{S+\mu S_\z}\p{-1+\mu S_\z}^m S_\bmu^n
@@
 \begin{align}
-  \D\p{\sigma_\mu; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
+  \D\p{s_\mu; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
   &= S_\mu^\ell \p{S - \ell - n - \mu S_\z} \p{1+\mu S_\z}^m S_\bmu^n
@@
 \end{align}
-Decoherence via jump operators $\sigma_\bmu^{(j)}$ only couples
-operators $S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$ to operators
+Decoherence via jump operators $s_\bmu^{(j)}$ only couples operators
+$S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$ to operators
 $S_\mu^\ell \p{\mu S_\z}^{m'} S_\bmu^n$ with $m'\le m$.  Decoherence
-via jump operators $\sigma_\mu^{(j)}$, meanwhile, makes operators
+via jump operators $s_\mu^{(j)}$, meanwhile, makes operators
 $S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$ ``grow'' in $m$ through the last
@@
 \begin{align}
-  \D\p{\sigma_\z; \O}
-  = \sum_j\p{\sigma_\z^{(j)} \O \sigma_\z^{(j)}
-    - \f12\sp{\sigma_\z^{(j)} \sigma_\z^{(j)},\O}_+}
-  = \sum_j \sigma_\z^{(j)} \O \sigma_\z^{(j)} - 2 S \O.
+  \D\p{s_\z; \O}
+  = \sum_j\p{s_\z^{(j)} \O s_\z^{(j)}
+    - \f12\sp{s_\z^{(j)} s_\z^{(j)},\O}_+}
+  = \sum_j s_\z^{(j)} \O s_\z^{(j)} - \f12 S \O.
 \end{align}
@@
 \begin{align}
-  \D\p{\sigma_\z; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
-  = -2\p{\ell+n} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n
-  + 4 \ell n S_\mu^{\ell-1} \p{S + \mu S_\z}
+  \D\p{s_\z; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
+  = -\f12\p{\ell+n} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n
+  + \ell n S_\mu^{\ell-1} \p{S + \mu S_\z}
   \p{-1 + \mu S_\z}^m S_\bmu^{n-1}.
@@
   &&
-  \gamma \equiv \gamma_\z \sigma_\z
-  + \gamma_+ \sigma_+ + \gamma_- \sigma_-.
+  \gamma \equiv \gamma_\z s_\z + \gamma_+ s_+ + \gamma_- s_-.
   \label{eq:D_general_single}
@@
   \gamma^\dag \O \gamma
-  = \abs{\gamma_\z}^2 \sigma_\z \O \sigma_\z
-  + \sum_\mu \p{\abs{\gamma_\mu}^2 \sigma_\bmu \O \sigma_\mu
-    + \gamma_\bmu^* \gamma_\mu \sigma_\mu \O \sigma_\mu
-    + \gamma_\z^* \gamma_\mu \sigma_\z \O \sigma_\mu
-    + \gamma_\bmu^* \gamma_\z \sigma_\mu \O \sigma_\z},
+  = \abs{\gamma_\z}^2 s_\z \O s_\z
+  + \sum_\mu \p{\abs{\gamma_\mu}^2 s_\bmu \O s_\mu
+    + \gamma_\bmu^* \gamma_\mu s_\mu \O s_\mu
+    + \gamma_\z^* \gamma_\mu s_\z \O s_\mu
+    + \gamma_\bmu^* \gamma_\z s_\mu \O s_\z},
 \end{align}
@@
   \gamma^\dag \gamma
-  = \abs{\gamma_\z}^2
-  + \sum_\mu \sp{\abs{\gamma_\mu}^2 \f12 \p{1-\mu\sigma_\z}
-    + \mu \p{\gamma_\z^*\gamma_\mu - \gamma_\bmu^*\gamma_\z}
-    \sigma_\mu},
+  = \f14 \abs{\gamma_\z}^2
+  + \f12 \sum_\mu \sp{\abs{\gamma_\mu}^2 \p{1-2\mu s_\z}
+    + \mu \p{\gamma_\z^*\gamma_\mu - \gamma_\bmu^*\gamma_\z} s_\mu},
 \end{align}
@@
   \D\p{\gamma;\O}
-  &= \sum_{X\in\set{\z,+,-}} \abs{\gamma_X}^2 \D\p{\sigma_X;\O}
+  &= \sum_{X\in\set{\z,+,-}} \abs{\gamma_X}^2 \D\p{s_X;\O}
   + \sum_{\mu,j}
-  \p{\gamma_\bmu^* \gamma_\mu \sigma_\mu^{(j)} \O \sigma_\mu^{(j)}
-    + \gamma_\z^* \gamma_\mu \sigma_\z^{(j)} \O \sigma_\mu^{(j)}
-    + \gamma_\bmu^* \gamma_\z \sigma_\mu^{(j)} \O \sigma_\z^{(j)}}
+  \p{\gamma_\bmu^* \gamma_\mu s_\mu^{(j)} \O s_\mu^{(j)}
+    + \gamma_\z^* \gamma_\mu s_\z^{(j)} \O s_\mu^{(j)}
+    + \gamma_\bmu^* \gamma_\z s_\mu^{(j)} \O s_\z^{(j)}}
   \notag \\
-  &\qquad -\f12 \sum_\mu \mu
+  &\qquad -\f14 \sum_\mu \mu
   \p{\gamma_\z^*\gamma_\mu - \gamma_\bmu^*\gamma_\z} \sp{S_\mu, \O}_+.
@@
   \sp{S_\mu, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
-  = S_\mu^{\ell+1} \sp{S_\z^m+\p{1+\mu S_\z}^m} S_\bmu^n
+  = S_\mu^{\ell+1} \sp{\p{\mu S_\z}^m+\p{1+\mu S_\z}^m} S_\bmu^n
   - n S_\mu^\ell \p{n-1+2\mu S_\z} \p{\mu S_\z}^m S_\bmu^{n-1}.
@@
 \begin{align}
-  \sum_j \sigma_\z^{(j)} S_\mu^\ell
-  \p{\mu S_\z}^m S_\bmu^n \sigma_\mu^{(j)}
-  - \f{\mu}{2} \sp{S_\mu, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
+  \sum_j s_\z^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\mu^{(j)}
+  - \f14 \mu \sp{S_\mu, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
   = K_{\ell mn;\mu} + L_{\ell mn;\mu}
@@
 \begin{align}
-  \sum_j \sigma_\mu^{(j)} S_\mu^\ell
-  \p{\mu S_\z}^m S_\bmu^n \sigma_\z^{(j)}
-  + \f{\mu}{2} \sp{S_\mu, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
+  \sum_j s_\mu^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\z^{(j)}
+  + \f14 \mu \sp{S_\mu, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
   = K_{\ell mn;\mu} + M_{\ell mn;\mu}
@@
   K_{\ell mn;\mu}
-  &\equiv \f{\mu}{2} S_\mu^{\ell+1}
+  &\equiv \f14 \mu S_\mu^{\ell+1}
   \sp{\p{1+\mu S_\z}^m-\p{\mu S_\z}^m} S_\bmu^n, \\
   L_{\ell mn;\mu}
-  &\equiv \mu n S_\mu^\ell \sp{-2S+2\ell+\f32\p{n-1}+\mu S_\z}
+  &\equiv -\mu n S_\mu^\ell \sp{S-\ell-\f34\p{n-1}-\f12\mu S_\z}
   \p{\mu S_\z}^m S_\bmu^{n-1} \notag \\
-  &\qquad - 2\mu\ell n\p{n-1} S_\mu^{\ell-1}
+  &\qquad - \mu\ell n\p{n-1} S_\mu^{\ell-1}
   \p{S+\mu S_\z} \p{-1+\mu S_\z}^m S_\bmu^{n-2}, \\
   M_{\ell mn;\mu}
-  &\equiv \mu n S_\mu^\ell \sp{2\p{S+\mu S_\z}\p{-1+\mu S_\z}^m
-    - \p{\f12\sp{n-1}+\mu S_\z}\p{\mu S_\z}^m} S_\bmu^{n-1}.
+  &\equiv \mu n S_\mu^\ell \sp{\p{S+\mu S_\z}\p{-1+\mu S_\z}^m
+    - \f12\p{\f12\sp{n-1}+\mu S_\z}\p{\mu S_\z}^m} S_\bmu^{n-1}.
 \end{align}
@@
   P_{\ell mn;\mu}
-  &\equiv \sum_j \sigma_\mu^{(j)} S_\mu^\ell
-  \p{\mu S_\z}^m S_\bmu^n \sigma_\mu^{(j)} \\
+  &\equiv \sum_j s_\mu^{(j)} S_\mu^\ell
+  \p{\mu S_\z}^m S_\bmu^n s_\mu^{(j)} \\
   &= n S_\mu^{\ell+1} \p{\mu S_\z}^m S_\bmu^{n-1}
@@
   = \sum_{X\in\set{\z,+,-}} \abs{\gamma_X}^2
-  \D\p{\sigma_X; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
+  \D\p{s_X; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
   + Q_{\ell mn;\mu}^{(\gamma)} + \sp{Q_{nm\ell;\mu}^{(\gamma)}}^\dag.
@@
 \end{align}
-we can say
+we can define the state
+\begin{align}
+  \ket{\phi} \equiv e^{-i\phi S_\z}\ket{\X},
+  &&
+  \p{\cos\phi S_\x + \sin\phi S_\y} \ket\phi = S \ket\phi,
+\end{align}
+and say that
 \begin{align}
   \bk{\phi|S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n|\phi}
-  = e^{i\mu\p{\ell-n}\phi} \bk{\X|S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n|\X},
-\end{align}
-where
-\begin{align}
-  \ket{\phi} \equiv e^{-i\phi S_\z}\ket{\X},
-  &&
-  \p{\cos\phi S_\x + \sin\phi S_\y} \ket\phi = S \ket\phi.
+  = e^{i\mu\p{\ell-n}\phi} \bk{\X|S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n|\X}.
 \end{align}
@@
   &&
-  P_{\v j} \equiv \prod_{\mu,a} \sigma_\mu^{\p{j^\mu_a}},
+  P_{\v j} \equiv \prod_{\mu,a} s_\mu^{\p{j^\mu_a}},
 \end{align}
 where $\v m \equiv \p{m_\z,m_+,m_-}\in\mathbb{N}_0^3$ with $m_\mu$ the
-number of $\sigma_\mu$ operators in each term of $\S_{\v m}$; $\v j$
-is an ordered list of indices for the spins addressed by $P_{\v j}$;
-and the sum over $\p{\v j;\v m}$ denotes a sum over all possible
-$\v j$ for which all indices in $\v j$ are distinct and
+number of $s_\mu$ operators in each term of $\S_{\v m}$; $\v j$ is an
+ordered list of indices for the spins addressed by $P_{\v j}$; and the
+sum over $\p{\v j;\v m}$ denotes a sum over all possible $\v j$ for
+which all indices in $\v j$ are distinct and
 $\abs{\set{j^\mu_a}} = m_\mu$, such that $j^\mu_a$ indexes the $a$-th
-spin addressed by a $\sigma_\mu$ operator in $P_{\v j}$.
+spin addressed by a $s_\mu$ operator in $P_{\v j}$.
 
@@
 \end{align}
-Collecting terms in which $r_{\mu\nu}$ of the $\sigma_\mu$ operators
-in $P_{\v j}$ address the same spin as a $\sigma_\nu$ operator in
-$P_{\v k}$, we have
+Collecting terms in which $r_{\mu\nu}$ of the $s_\mu$ operators in
+$P_{\v j}$ address the same spin as a $s_\nu$ operator in $P_{\v k}$,
+we have
 \begin{align}
@@
   Q_{\v K} \equiv \prod_{\mu,\nu,a}
-  \sigma_\mu^{(K^{\mu\nu}_a)} \sigma_\nu^{(K^{\mu\nu}_a)}.
+  s_\mu^{(K^{\mu\nu}_a)} s_\nu^{(K^{\mu\nu}_a)}.
 \end{align}
@@
   = \prod_{\mu,\nu} \prod_{a=1}^{r_{\mu\nu}}
-  \sigma_\mu^{(K^{\mu\nu}_a)} \sigma_\nu^{(K^{\mu\nu}_a)}
-  \simeq \bigotimes_{\mu,\nu} \bigotimes_{a=1}^{r_{\mu\nu}}
-  \sigma_\mu \sigma_\nu
+  s_\mu^{(K^{\mu\nu}_a)} s_\nu^{(K^{\mu\nu}_a)}
+  \simeq \bigotimes_{\mu,\nu} \bigotimes_{a=1}^{r_{\mu\nu}} s_\mu s_\nu
   = \bigotimes_{\mu,\nu} \bigotimes_{a=1}^{r_{\mu\nu}}
-  \sum_\rho \eta_{\mu\nu\rho} \sigma_\rho,
+  \sum_\rho \eta_{\mu\nu\rho} s_\rho,
   \label{eq:Q_K_eta}
@@
 $\eta_{\mu\nu\rho}$ is defined by
-$\sigma_\mu\sigma_\nu=\sum_\rho\eta_{\mu\nu\rho}\sigma_\rho$.  Unlike
-the sums over $\mu,\nu$ in most of the work above, the sum over $\rho$
-here includes an index for the identity operator $\sigma_0\equiv\1$.
-Distributing the product of sums into a sum of products gives us
+$s_\mu s_\nu=\sum_\rho\eta_{\mu\nu\rho}s_\rho$.  Unlike the sums over
+$\mu,\nu$ in most of the work above, the sum over $\rho$ here includes
+an index for the identity operator $s_0\equiv\1$.  Distributing the
+product of sums into a sum of products gives us
 \begin{align}
@@
   \simeq \sum_{\v\rho} \bigotimes_{\mu,\nu,a}
-  \eta_{\mu\nu\rho^{\mu\nu}_a} \sigma_{\rho^{\mu\nu}_a}
+  \eta_{\mu\nu\rho^{\mu\nu}_a} s_{\rho^{\mu\nu}_a}
   = \sum_{\v\rho} \sp{\prod_{\mu,\nu,a}
     \p{\eta_{\mu\nu\rho^{\mu\nu}_a}}^{\rho^{\mu\nu}_a}}
-  \bigotimes_{\mu,\nu,a} \sigma_{\rho^{\mu\nu}_a},
+  \bigotimes_{\mu,\nu,a} s_{\rho^{\mu\nu}_a},
   \label{eq:Q_K_rho}
@@
   \bigotimes_{\kappa}
-  \sigma_\kappa^{\otimes\sum_{\mu,\nu}\tilde\rho^{\mu\nu}_\kappa}
+  s_\kappa^{\otimes\sum_{\mu,\nu}\tilde\rho^{\mu\nu}_\kappa}
   \simeq \sum_{\tilde{\v\rho}} g_{\tilde{\v\rho}} \v\eta_{\tilde{\v\rho}}
@@
   \otimes \bigotimes_{\kappa\ne0}
-  \sigma_\kappa^{\otimes\sum_{\mu,\nu}\tilde\rho^{\mu\nu}_\kappa},
+  s_\kappa^{\otimes\sum_{\mu,\nu}\tilde\rho^{\mu\nu}_\kappa},
   \label{eq:Q_K_rho_simp}
@@
 \end{align}
-where the sum over $\p{\v\rho;\v r}$ denotes a sum over all
-values of $\v\rho$ and with the restrictions
-\begin{align}
-  \abs{\set{\rho^{\mu\nu}_\kappa}} = \abs{\set{\sigma_\kappa}},
+where the sum over $\p{\v\rho;\v r}$ denotes a sum over all values of
+$\v\rho$ and with the restrictions
+\begin{align}
+  \abs{\set{\rho^{\mu\nu}_\kappa}} = \abs{\set{s_\kappa}},
   &&
@@
 permutation of indices involving the identity operators in
-\eqref{eq:Q_K_rho_simp}; and
-the number of $\sigma_\mu$ operators in each term of
-$\tilde\S_{\v\ell_{\v m\v n\v r\v\rho}}$ is
+\eqref{eq:Q_K_rho_simp}; and the number of $s_\mu$ operators in each
+term of $\tilde\S_{\v\ell_{\v m\v n\v r\v\rho}}$ is
 \begin{align}
@@
 counts the number of operators addressed by
-$\sigma_\mu\sigma_\nu=\sum_\kappa\eta_{\mu\nu\kappa}\sigma_\kappa$, if
-$\eta_{\mu\nu\kappa}=0$ for all $\kappa$,
-i.e.~$\sigma_\mu\sigma_\nu=0$, then any term with $r_{\mu\nu}>0$ will
-vanish.  Our second restriction on the sums in \eqref{eq:SS_simp} is
-therefore to fix $r_{\mu\nu}=0$ for all $\mu,\nu$ with
-$\sigma_\mu\sigma_\nu=0$.  Finally, we can avoid the nullifying effect
-of any additional zeros in the multiplication table
+$s_\mu s_\nu=\sum_\kappa\eta_{\mu\nu\kappa} s_\kappa$, if
+$\eta_{\mu\nu\kappa}=0$ for all $\kappa$, i.e.~$s_\mu s_\nu=0$, then
+any term with $r_{\mu\nu}>0$ will vanish.  Our second restriction on
+the sums in \eqref{eq:SS_simp} is therefore to fix $r_{\mu\nu}=0$ for
+all $\mu,\nu$ with $s_\mu s_\nu=0$.  Finally, we can avoid the
+nullifying effect of any additional zeros in the multiplication table
 $\eta_{\mu\nu\kappa}$, which appear as factors in

### 2018-11-08 soc_squeezing/scripts/test_correlators.py

@@ -29,8 +29,9 @@ ivp_tolerance = 1e-5
 def rand(magnitude = 1):
     return ( random() + 1j*random() ) * magnitude
 
-dec_rates = [ (rand(10), rand(10), rand(10)),
-              (rand(10), rand(10), rand(10)) ]
+mag = 5
+dec_rates = [ (rand(mag), rand(mag), rand(mag)),
+              (rand(mag), rand(mag), rand(mag)) ]
 dec_mat = np.array([ [ random() for jj in range(3) ] for kk in range(3) ])
 
 h_rand = { (ll,mm,nn) : random()
@@ -49,13 +50,13 @@ s_z_j = []
 s_p_j = []
 s_m_j = []
 for jj in range(N):
-    s_z_j.append(qt.tensor([ I2 ] * jj + [ qt.sigmaz() ] + [ I2 ] * (N-jj-1)))
+    s_z_j.append(qt.tensor([ I2 ] * jj + [ qt.sigmaz()/2 ] + [ I2 ] * (N-jj-1)))
     s_p_j.append(qt.tensor([ I2 ] * jj + [ qt.sigmap() ] + [ I2 ] * (N-jj-1)))
     s_m_j.append(qt.tensor([ I2 ] * jj + [ qt.sigmam() ] + [ I2 ] * (N-jj-1)))
 
 S_p = sum(s_p_j)
 S_m = sum(s_m_j)
-S_z = sum(s_z_j) / 2
+S_z = sum(s_z_j)
 S_x =   1/2 * ( S_p + S_m )
 S_y = -1j/2 * ( S_p - S_m )
 

### 2018-11-08 soc_squeezing/scripts/test_decoherence.py

@@ -17,13 +17,13 @@ s_z_j = []
 s_p_j = []
 s_m_j = []
 for jj in range(N):
-    s_z_j.append(qt.tensor([ I2 ] * jj + [ qt.sigmaz() ] + [ I2 ] * (N-jj-1)))
+    s_z_j.append(qt.tensor([ I2 ] * jj + [ qt.sigmaz()/2 ] + [ I2 ] * (N-jj-1)))
     s_p_j.append(qt.tensor([ I2 ] * jj + [ qt.sigmap() ] + [ I2 ] * (N-jj-1)))
     s_m_j.append(qt.tensor([ I2 ] * jj + [ qt.sigmam() ] + [ I2 ] * (N-jj-1)))
 
 S_p = sum(s_p_j)
 S_m = sum(s_m_j)
-S_z = sum(s_z_j) / 2
+S_z = sum(s_z_j)
 
 def acomm(A,B):
     return A*B + B*A
@@ -88,27 +88,27 @@ for l, m, n in itertools.product(range(N), repeat = 3):
 
         ##############################
 
-        d_z = -2*(l+n) * S_mu**l * (mu*S_z)**m * S_nu**n
+        d_z = -1/2*(l+n) * S_mu**l * (mu*S_z)**m * S_nu**n
         if l >= 1 and n >= 1:
-            d_z += 4*l*n * S_mu**(l-1) * (S+mu*S_z) * (-1+mu*S_z)**m * S_nu**(n-1)
+            d_z += l*n * S_mu**(l-1) * (S+mu*S_z) * (-1+mu*S_z)**m * S_nu**(n-1)
 
         ##############################
 
         def K(l,m,n,mu):
-            return mu/2 * S_mu**(l+1) * ( (1+mu*S_z)**m - (mu*S_z)**m ) * S_nu**n
+            return mu/4 * S_mu**(l+1) * ( (1+mu*S_z)**m - (mu*S_z)**m ) * S_nu**n
 
         def L(l,m,n,mu):
             L = 0*II
             if n >= 1:
-                L += mu*n * S_mu**l * (-2*S+2*l+3/2*(n-1)+mu*S_z) * (mu*S_z)**m * S_nu**(n-1)
+                L += -mu*n * S_mu**l * (S-l-3/4*(n-1)-mu/2*S_z) * (mu*S_z)**m * S_nu**(n-1)
             if n >= 2 and l >= 1:
[…]

### 2018-11-08 soc_squeezing/scripts/test_product_factored.py

@@ -11,7 +11,7 @@ N = 3
 op_cap = 3
 
 I2 = qt.qeye(2)
-sz = qt.sigmaz()
+sz = qt.sigmaz()/2
 sp = qt.sigmap()
 sm = qt.sigmam()
 II = qt.tensor([ I2 ] * N)
@@ -61,8 +61,10 @@ def poch(nn, kk, vals = {}):
     return np.prod([ nn - cc for cc in range(kk) ])
 
 def eta_val(mu, nu, rho):
-    fac = 2 if spin_op(rho) in [ I2, sz ] else 1
-    return ( spin_op(mu)*spin_op(nu) * spin_op(rho).dag() ).tr() / fac
+    if spin_op(rho) == I2: fac = 1/2
+    elif spin_op(rho) == sz: fac = 2
+    else: fac = 1
+    return ( spin_op(mu)*spin_op(nu) * spin_op(rho).dag() ).tr() * fac
 
 eta = np.array([ [ [ eta_val(mu,nu,kk) for kk in range(4) ]
                    for nu in range(3) ]

### 2018-11-16 soc_squeezing/cumulant_expansions.tex
modified 'cumulant expansion' text into the outline of a paper

@@
-\documentclass[aps,notitlepage,nofootinbib,11pt]{revtex4-1}
+\documentclass[aps,11pt,notitlepage,nofootinbib]{revtex4-1}
 
@@
 
-\title{Short-time simulation of collective spin Hamiltonians with
-  arbitrary decoherence}
+\title{Short-time simulation of collective quantum spin systems with
+  decoherence}
 
 \author{Michael A. Perlin}
+\email{mika.perlin@gmail.com}
+\author{Ana Maria Rey}
+\affiliation{JILA, National Institute of Standards and Technology and
+  University of Colorado, 440 UCB, Boulder, Colorado 80309, USA}
+\affiliation{Center for Theory of Quantum Matter, 440 UCB, Boulder,
+  Colorado 80309, USA}
+\affiliation{Department of Physics, University of Colorado, 390 UCB,
+  Boulder, Colorado 80309, USA}
+
+\begin{abstract}
+  Abstract goes here.
+\end{abstract}
 
@@
 
-\tableofcontents
-
-\newpage
-
 \section{Introduction}
 
-Previous work has shown how to realize spin-squeezing Hamiltonians
-with a variety of ultracold atomic and trapped ion systems.  While
-simulating the dynamics of initially Gaussian states evolving under
-these Hamiltonians is straightforward in the context of a closed
-quantum system, here we develop a method to simulate short-time
-dynamics in the presence of decoherence.  In particular, we will
-consider dynamics induced by the one-axis twisting (OAT), two-axis
-twisting (TAT), and transverse field (TVF) Hamiltonians of the form
+Collective spin systems attract considerable attention for providing a
+window into many-body quantum phenomena that are both experimentally
+accessible\cite{fernholz2008spin, takano2009spin,
+  riedel2010atomchipbased, gross2010nonlinear, martin2013quantum,
+  bohnet2016quantum, hosten2016quantum, norcia2018cavitymediated} and
+metrologically useful\cite{wineland1992spin, kitagawa1993squeezed,
+  ma2011quantum}.  Theoretical interest in such systems dates back to
+the mid-twentieth century with the introduction of the
+Lipkin-Meshkov-Glick (LMG) model as a toy model for testing many-body
+approximation methods in contemporary nuclear
+physics\cite{lipkin1965validity, meshkov1965validity,
+  glick1965validity}.  Since then, collective system systems have
+additionally featured in theoretical studies of quantum
+criticality\cite{latorre2005entanglement, alcalde2007functional,
+  morrison2008collective, sarandy2009classical, wang2012quantum,
+  majd2014lmg}, non-equilibrium
+phenomena\cite{morrison2008dissipationdriven,
+  bhattacherjee2014nonequilibrium, klinder2015dynamical,
+  zhiqiang2017nonequilibrium, smale2018observation}, and most notably
+spin squeezing\cite{wineland1992spin, kitagawa1993squeezed,
+  ma2011quantum}.  With the development of advanced atomic trapping,
+cooling, and control techniques, collective spin systems have been
+experimentally realized in atomic ensembles\cite{fernholz2008spin,
+  takano2009spin}, Bose-Einstein
+condensates\cite{riedel2010atomchipbased, gross2010nonlinear},
+ultracold Fermi gasses\cite{martin2013quantum}, trapped
+ions\cite{bohnet2016quantum}, and optical
+cavities\cite{hosten2016quantum, norcia2018cavitymediated}.
+
+One of the primary current motivations for studying collective spin
+systems is their capability to dynamically generate spin-squeezed
+states, which enable measurement of phase angles $\phi$ with an error
+$\Delta\phi$ below the standard quantum limit
+$\Delta\phi\sim1/\sqrt{N}$ set by quantum projection noise from the
+measurement of $N$ independent particles\cite{ma2011quantum}.  The
+nonlinear dynamics necessary to generate spin-squeezed states can be
+implemented with long-range collisional, photon-, or phonon-mediated
+interactions, typically resulting in a one-axis twising (OAT) model
+described by the Hamiltonian\cite{kitagawa1993squeezed}
 \begin{align}
   H_{\t{OAT}} = \chi S_\z^2,
-  &&
-  H_{\t{TAT}}^{\z,\y} = \chi \p{S_\z^2 - S_\y^2},
-  &&
-  H_{\t{TVF}} = \chi S_\z^2 + \Omega S_\x.
-\end{align}
-In order to determine the squeezing behavior of these Hamiltonians, we
-will study the dynamics of Heisenburg operators $\O$, whose evolution
-under a Hamiltonian $H$ is given by
-\begin{align}
-  \f{d}{dt} \O = i\sp{H,\O}_- + \sum_\J \gamma_\J \D\p{\J;\O},
-  &&
+  \label{eq:OAT}
+\end{align}
+where $S_\z\equiv\sum_{j=1}^N\sigma_\z^{(j)}/2$ is a collective
+spin-$z$ operator defined in terms of Pauli-$z$ operators
+$\sigma_\z^{(j)}$ addressing spin $j$.  The OAT model prepares a
+spin-squeezed state that allows measurement of a phase angle $\phi$
+with error $\Delta\phi\sim N^{-2/3}$ at times
+$t\sim\chi^{-1}N^{-2/3}$\cite{ma2011quantum}.  Although a truly
+collective spin model requires uniform, all-to-all interactions, as
+long as measurements do not distinguish between constituent particles
+even non-uniform systems can be effectively described by a uniform
+model with renormalized parameters\cite{hu2015entangled}.  Various
+works have proposed improvements over OAT through the addition of
+auxiliary control fields\cite{law2001coherent, liu2011spin,
+  huang2015twoaxis, muessel2015twistandturn} or a modified
+vacuum\cite{hu2017vacuum}, thereby realizing different collective spin
+models.
+
+While the OAT model admits exact
+solutions\cite{foss-feig2013nonequilibrium}, the same is not true of
+more general, non-Ising Hamiltonians.  Nonetheless, in the absence of
+decoherence, permutation symmetry and total spin conservation divide
+the total Hilbert space of a collective spin system into superseletion
+sectors that grow only linearly with system size, thereby admitting
+efficient classical simulation of its dynamics.  If decoherence is
+present, but sufficiently weak, dynamics of collective spin systems
+are still numerically solvable via ``quantum jump'' Monte Carlo
+methods\cite{plenio1998quantumjump, zhang2018montecarlo} (also known
+as ``quantum trajectory'' or ``Monte Carlo wavefunction'' methods)
+that can reproduce all expectation values of interest.  When
+decoherence is strong, however, or when the number of jump operators
+grows extensively with system size (e.g.~for single-spin decay), these
+Monte Carlo methods can take a prohibitively long time to converge, as
+simulations become dominated by incoherent jumps that generate large
+numbers of distinct trajectories over which to average.
+
+In this work, we develop a new method to simulate short-time dynamics
+of collective spin systems with arbitrary-strength
+permutationally-symmetric decoherence.  This method is based on a
+short-time expansion of exact solutions to the equations of motion for
+Heisenberg operators.  Evaluating this expansion requires use of the
+structure factors of a collective spin operator algebra; the
+calculation of these structure factors is one of the main technical
+results of this work.
+
+\section{Theory}
+
+In this section we provide the basic theory for our method to compute
+expectation values of collective spin operators, deferring lengthy
+derivations to the Appendices.  We consider systems composed of $N$
+distinct spin-1/2 particles.  Defining individual spin-1/2 operators
+$s_{n=\x,\y,\z}\equiv\sigma_n/2$ and $s_\pm\equiv s_\x\pm is_\y$ with
+Pauli operators $\sigma_n$, we denote an operator which acts with
+$s_n$ on the spin indexed by $j$ and trivially (i.e.~with the identity
+$\1$) on all other spins by $s_n^{(j)}$.  We then define the
+collective spin operators $S_n\equiv\sum_{j=1}^Ns_n^{(j)}$ for
+$n\in\set{\x,\y,\z,+,-}$.  Identifying the set $\set{\S_{\v m}}$ as a
+basis for all collective spin operators, with
+$\v m\equiv\p{m_+,m_\z,m_-}\in\N_0^3$ and
+$\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$, we can expand all
+collective spin Hamiltonians in the form
+\begin{align}
+  H = \sum_{\v m} h_{\v m} \S_{\v m},
+  \label{eq:H_general}
+\end{align}
+where Hermiticity requires that $h_{m_+,m_\z,m_-}=h_{m_-,m_\z,m_+}^*$.
+The time evolution of a particular Heisenberg operator $\S_{\v m}$
+under a Hamiltonian of the form in Eq.~\eqref{eq:H_general} is then
+given by
+\begin{align}
+  \f{d}{dt}\S_{\v m}
+  = i\sum_{\v n}h_{\v n}\sp{\S_{\v n},\S_{\v m}}_-
+  + \sum_\J \gamma_\J \D\p{\J;\S_{\v m}}
+  \equiv \sum_{\v n} T_{\v m\v n} \S_{\v n},
+  \label{eq:EOM}
+\end{align}
+where $\sp{X,Y}_\pm\equiv XY\pm YX$, $\J$ is a set of jump operators
+with a corresponding decoherence rate $\gamma_\J$, and
+\begin{align}
   \D\p{\J;\O}
-  \equiv \sum_{J\in\J}\p{J^\dag \O J - \f12\sp{J^\dag J,\O}_+},
-  \label{eq:EOM}
-\end{align}
-where $\J$ is a set of jump operators with a corresponding decoherence
-rate $\gamma_\J$ and $\sp{X,Y}_\pm\equiv XY\pm YX$.  Our measure of
-spin squeezing is
+  \equiv \sum_{J\in\J} \p{J^\dag \O J - \f12\sp{J^\dag J,\O}_+}.
+\end{align}
+The commutator in Eq.~\eqref{eq:H_general} can be determined from the
+structure factors of $f_{\v n\v m\v\ell}$ of the general product
+$\S_{\v n}\S_{\v m}=\sum_{\v\ell}f_{\v n\v m\v\ell}S_{\v\ell}$ worked
+out in Appendix \ref{sec:prod_general}, while the effects of
+decoherence from jump operators (i.e.~elements of $\J$) of the form
+$\gamma^{(j)}=\sum_n\gamma_ns_n^{(j)}$ and $\Gamma=\sum_n\Gamma_nS_n$
+are respectively worked out in Appendices \ref{sec:decoherence_single}
+and \ref{sec:decoherence_collective}.  These ingredients are
+sufficient to compute matrix elements $T_{\v m\v n}$ of the time
+derivative operator $d/dt$ in Eq.~\eqref{eq:EOM} in most cases of
+interest.
+
+The time derivative operator $d/dt$ will generally couple spin
+operators $\S_{\v m}$ to spin operators $\S_{\v n}$ with higher
+``weight'', i.e.~with $\abs{\v n}>\abs{\v m}$, where
+$\abs{\v\ell}\equiv\sum_\lambda\ell_\lambda$.  The growth of operator
+weight signifies the growth of many-body correlations.  In practice,
+keeping track of this growth will eventually require more
+computational resources than are available, meaning we must somehow
+truncate our equations of motion.  The simplest truncation strategy
+would be to take
+\begin{align}
+  \f{d}{dt} \S_{\v m}
+  \to \sum_{w\p{\v n}<W} T_{\v m\v n} \S_{\v n}
+  \label{eq:weight_truncation}
+\end{align}
+for some weight measure $w$, e.g.~$w\p{\v n}=\abs{\v n}$, and cutoff
+$W$.  The truncation in \eqref{eq:weight_truncation} will close the
+system of differential equations defined by \eqref{eq:EOM}, and allow
+us to solve it using standard numerical methods.  Initial conditions
+for this system of differential equations, i.e.~expectation values of
+collective spin operators with respect to states that are simple to
+prepare experimentally, are provided in Appendix
+\ref{sec:initial_conditions}.
+
+The truncation strategy in Eq.~\eqref{eq:weight_truncation} has a few
+limitations:
+\begin{enumerate*}[label=(\roman*)]
+\item simulating a system of differential equations for a large number
+  of operators can be quite time-consuming,
+\item the weight measure $w$ may need to be chosen carefully, as the
+  optimal measure is generally system-dependent,
+\item simulation results can only be trusted up to the time at which
+  the initial value of operators $\S_{\v m}$ with weight
+  $w\p{\v m}\ge W$ have a non-negligible contribution to operators of
+  interest.\label{pt:limitation}
+\end{enumerate*}
+The last limitation in particular unavoidably applies in some form to
+any method tracking only a subset of all relevant operators.  We
+therefore devise an alternate truncation strategy built around point
+\ref{pt:limitation}.
+
+We can expand Heisenberg operators in a Taylor series about $t=0$ to
+write
+\begin{align}
+  \bk{\S_{\v m}}
+  = \sum_{k\ge0} \f{t^k}{k!} \bk{\f{d^k}{dt^k} \S_{\v m}}_{t=0}
+  = \sum_{k\ge0} \f{t^k}{k!}
+  \sum_{\v n} T_{\v m\v n;k} \bk{\S_{\v n}}_{t=0},
+  \label{eq:time_series}
+\end{align}
+where the matrix elements of the $k$-th time derivative are given by
+\begin{align}
+  T_{\v m\v n;0} \equiv \delta_{\v m\v n},
+  &&
+  T_{\v m\v n;1} \equiv T_{\v m\v n},
+  &&
+  T_{\v m\v n;k>1}
+  \equiv \sum_{\v p_1,\v p_2,\cdots,\v p_{k-1}}
+  T_{\v m\v p_1} T_{\v p_1\v p_2} T_{\v p_2\v p_3}
+  \cdots T_{\v p_{k-1}\v n},
+\end{align}
+where $\delta_{\v m\v n}=1$ if $\v m=\v n$ and 0 otherwise.  For
+sufficiently short times, we can truncate the series in
+Eq.~\eqref{eq:time_series} by taking
+\begin{align}
+  \bk{\S_{\v m}}
+  \to \sum_{k=0}^M \f{t^k}{k!}
+  \sum_{\v n} T_{\v m\v n;k} \bk{\S_{\v n}}_{t=0}.
+\end{align}
+Unlike the truncation strategy in Eq.~\eqref{eq:weight_truncation},
+the nonzero entries $T_{\v m\v n;k}$ for $k=0,1,\cdots,M$ now
+explicitly tell us which operators are relevant for computing the
+expectation value $\bk{\S_{\v m}}$ to a fixed order $M$.  Furthermore,
+after computing all necessary matrix elements $T_{\v m\v n;k}$ and
+expectation values $\bk{\S_{\v n}}_{t=0}$, computing $\bk{\S_{\v m}}$
+takes only an $\O\p{M}$ amount of computer time independent of the
+system time $t$.  Finally, we note that the relation
+$\S_{\v m}^\dag=\S_{\v m^*}$ for $\v m^*\equiv\p{m_-,m_\z,m_+}$ can be
+exploited to cut both the number of matrix elements $T_{\v m\v n}$ and
+expectation values $\bk{\S_{\v m}}_{t=0}$ that we need to explicitly
+compute roughly in half.
+
+\section{Benchmarking}
+
+Our measure of spin squeezing is
 \begin{align}
@@
 This squeezing parameter is entirely determined by one- and two-spin
-correlators of the form $\bk{S_\alpha}$ and $\bk{S_\alpha S_\beta}$;
-our task is therefore to compute these expectation values in all
-scenarios of interest.
-
-\section{One-axis twisting}
-
-The simplest case is that of one-axis twisting, as we can write the
-Hamiltonian in the form
-\begin{align}
-  H_{\t{OAT}} = \f14 \chi \sum_{j,k} \sigma_\z^{(j)} \sigma_\z^{(k)}
+correlators of the form $\bk{S_\alpha}$ and $\bk{S_\alpha S_\beta}$.
+
+\section{Discussion}
+
+Many-body correlators and cumulants \`a la
+Ref.~\cite{schweigler2017solving}?
+
+\section{Conclusion}
+
+\bibliography{\jobname}
+
+\newpage
+\appendix
+
+\section{Exact results for the one-axis twisting model}
+\label{sec:OAT}
+
+The one-axis twisting (OAT) Hamiltonian for $N$ spin-1/2 particles
+takes the form
+\begin{align}
+  H_{\t{OAT}}
+  = \chi S_\z^2
+  = \f14 \chi \sum_{j,k} \sigma_\z^{(j)} \sigma_\z^{(k)}
   = \f12 \chi \sum_{j<k} \sigma_\z^{(j)} \sigma_\z^{(k)} + \f14 N \chi,
 \end{align}
-which admits an exact solution previously worked out by Foss-Feig et
-al.~in ref.~\cite{foss-feig2013nonequilibrium}.  Adapting exact
-expectation values with respect to the initial state $\ket{\X}$
-satisfying $S_\x\ket{\X}=S\ket{\X}$, in terms of $S\equiv N/2$ and
-$\mu,\nu\in\set{+1,-1}$ we find that
+where $\sigma_z^{(j)}$ represents a Pauli-$z$ operator acting on spin
+$j$.  This Hamiltonian admits an exact solution previously worked out
+in Ref.~[\citenum{foss-feig2013nonequilibrium}].  Adapting the
+expectation values therein with respect to the initial state
+$\ket{\X}$ satisfying $S_\x\ket{\X}=S\ket{\X}$, in terms of
+$S\equiv N/2$ and $\mu,\nu\in\set{+1,-1}$ we find that
 \begin{align}
@@
 \begin{align}
-  \gamma \equiv \f12 \p{\gamma_+ - \gamma_-},
+  \gamma \equiv -\f12 \p{\gamma_+ - \gamma_-},
   &&
@@
 \end{align}
-defined respectively in terms of uncorrelated spin dephasing,
-excitation, and decay rates $\gamma_\z$, $\gamma_+$, $\gamma_-$.  The
-last ingredients we need to compute spin squeezing at any time are
+defined respectively in terms of uncorrelated single-spin decay,
+excitation, and dephasing rates $\gamma_-$, $\gamma_+$, $\gamma_\z$
+(respectively $\Gamma_{\t{ud}}$, $\Gamma_{\t{du}}$, and
+$\Gamma_{\t{el}}$ in Ref.~[\citenum{foss-feig2013nonequilibrium}]).
+The last ingredients we need to compute spin squeezing at any time are
 $\bk{S_\z}$ and $\bk{S_\z^2}$.  As these operators commute with the
@@
 The expectation values in \eqref{eq:S+_OAT}-\eqref{eq:SS+-_OAT} and
-\eqref{eq:Sz_OAT} are sufficient to compute the squeezing parameter
-$\xi^2$ defined in \eqref{eq:squeezing} at any time throughout
-evolution of the initial state $\ket{\X}$ under $H_{\t{OAT}}$.
-
-\section{Short-time evolution via truncated equations of motion}
-
-Unlike the one-axis twisting Hamiltonian, the two-axis twisting and
-transverse-field Hamiltonians do not admit exact analytical solutions.
-We will therefore simulate dynamics under these Hamiltonians
-numerically using alternate methods.  The basic strategy will be to
-first find equations of motion for all operators of interest using
-some complete basis for collective-spin operators.  These equations of
-motion will generally yield a system of equations involving a number
-of operators that grows too fast with system size for exact
-simulation.  We will therefore look for ways to truncate the equations
-of motion to arrive at a reasonably-sized expressions that are
-practical to compute.  To benchmark this method, we test it against
-collective Hamiltonian dynamics in the absence of decoherence, exact
-one-axis twisting results in the presence of decoherence, and exact
-simulations of small systems.
-
-Ultimately, we are interested in the dynamics of collective spin
-operators.  Choosing a basis $\set{\S_{\v m}}$ for these operators,
-e.g.~$\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$ with
-$\v m\equiv\p{m_+,m_\z,m_-}\in\mathbb{N}_0^3$, we can expand all
-collective-spin Hamiltonians in the form
-\begin{align}
-  H = \sum_{\v m} h_{\v m} \S_{\v m},
-  \label{eq:H_general}
-\end{align}
-where if $\S_{\v m}=S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$ then Hermiticity
-requires that $h_{m_+,m_\z,m_-}=h_{m_-,m_\z,m_+}^*$.  The one-axis
-twisting Hamiltonian with a transverse field, for example, is
-\begin{align}
-  H_{\t{TVF}}
-  = \chi S_\z^2 + \Omega S_\x
-  = \chi S_\z^2 + \f{\Omega}{2} S_+ + \f{\Omega}{2} S_-,
-\end{align}
-while the two-axis twisting Hamiltonians about $\p{z,y}$ and $\p{y,x}$
-are
-\begin{align}
-  H_{\t{TAT}}^{\z,\y}
-  \propto S_\z^2 - S_\y^2
-  = S_\z^2 + \f14 S_+^2 + \f14 S_-^2
-  - \f12 S_+ S_- + \f12 S_\z,
-\end{align}
-and
-\begin{align}
-  H_{\t{TAT}}^{\y,\x}
-  \propto S_\y^2 - S_\x^2
-  = -\f12 S_+^2 - \f12 S_-^2.
-\end{align}
-The evolution of a general operator $\S_{\v m}$ under a Hamiltonian of
-the form in \eqref{eq:H_general} is given by
-\begin{align}
-  \f{d}{dt} \S_{\v m}
-  = i \sum_{\v n} h_{\v n}
-  \sp{\S_{\v n}, \S_{\v m}}_-
-  + \sum_\J \gamma_\J \D\p{\J; \S_{\v m}}
-  \equiv \sum_{\v n} T_{\v m\v n} \S_{\v n},
-  \label{eq:time_derivative}
-\end{align}
-where commutator and decoherence terms necessary for computing the
-matrix elements $T_{\v m\v n}$ of the time derivative operator $d/dt$
-are worked out Appendices \ref{sec:prod_general},
-\ref{sec:decoherence_single}, and \ref{sec:decoherence_collective}.
-
-The time derivative operator $d/dt$ will generally couple spin
-operators $\S_{\v m}$ to spin operators $\S_{\v n}$ with higher
-``weight'', i.e.~with $\abs{\v n}>\abs{\v m}$ where
-$\abs{\v\ell}\equiv\sum_\lambda\ell_\lambda$.  The growth of operator
-weight signifies the growth of many-body correlations.  In practice,
-keeping track of this growth will eventually require more
-computational resources than are available, so we truncate it by
-taking
-\begin{align}
-  \f{d}{dt} \S_{\v m}
-  = \sum_{\v n} T_{\v m\v n} \S_{\v n}
-  \to \sum_{w\p{\v n}<W} T_{\v m\v n} \S_{\v n}
-  \label{eq:time_derivative_truncation}
-\end{align}
-for some weight measure $w$, e.g.~$w\p{\v n}=\abs{\v n}$, and cutoff
-$W$.  The truncation in \eqref{eq:time_derivative_truncation} will
-close the system of differential equations defined by
-\eqref{eq:time_derivative}, and allow us to solve it using standard
-numerical methods.  The initial conditions for this system of
-differential equations, i.e.~expectation values of operators with
-respect to states that are simple to prepare, are provided in Appendix
-\ref{sec:initial_conditions}.
-
-The truncation strategy in \eqref{eq:time_derivative_truncation} has a
-few limitations:
-\begin{enumerate*}[label=(\roman*)]
-\item simulating a system of differential equations for a large number
-  of operators can be very time-consuming,
-\item the weight measure $w$ may need to be chosen carefully, as the
-  optimal measure is generally system-dependent,
-\item simulation results can only be trusted up to the time at which
-  the initial value of operators $\S_{\v m}$ with weight
-  $w\p{\v m}\ge W$ have a non-negligible contribution to operators of
-  interest.\label{pt:limitation}
-\end{enumerate*}
-The last limitation in particular {\it necessarily} applies in some
-form to any method tracking only a subset of all relevant operators.
-Embracing the unavoidable, we therefore try to devise an alternate
-truncation strategy built around \ref{pt:limitation}.
-
-We can expand Heisenberg operators in a Taylor series about $t=0$ to
-write
-\begin{align}
-  \bk{\S_{\v m}}
-  = \sum_{k\ge0} \f{t^k}{k!} \bk{\f{d^k}{dt^k} \S_{\v m}}_{t=0}
-  = \sum_{k\ge0} \f{t^k}{k!}
-  \sum_{\v n} T_{\v m\v n;k} \bk{\S_{\v n}}_{t=0},
-  \label{eq:time_series}
-\end{align}
-where the matrix elements of the $k$-th time derivative are given by
-\begin{align}
-  T_{\v m\v n;0} \equiv \delta_{\v m\v n},
-  &&
-  T_{\v m\v n;1} \equiv T_{\v m\v n},
-  &&
-  T_{\v m\v n;k>1}
-  \equiv \sum_{\v p_1,\v p_2,\cdots\v p_{k-1}}
-  T_{\v m\v p_1} T_{\v p_1\v p_2} T_{\v p_2\v p_3}
-  \cdots T_{\v p_{k-1}\v n}.
-\end{align}
-Instead of explicitly choosing a weight function that determines which
-operators to keep track of, we can now truncate the series in
-\eqref{eq:time_series} by taking
-\begin{align}
-  \bk{\S_{\v m}}
-  = \sum_{k\ge0} \f{t^k}{k!}
-  \sum_{\v n} T_{\v m\v n;k} \bk{\S_{\v n}}_{t=0}
-  \to \sum_{k=0}^M \f{t^k}{k!}
-  \sum_{\v n} T_{\v m\v n;k} \bk{\S_{\v n}}_{t=0}.
-\end{align}
-The nonzero entries of $T_{\v m\v n;k}$ for $k=0,1,\cdots,M$ now tell
-us which operators we need to keep track of in order to compute the
-expectation value $\bk{\S_{\v m}}$.  Furthermore, after computing all
-relevant matrix elements $T_{\v m\v n;k}$ and expectation values
-$\bk{\S_{\v n}}_{t=0}$, computing $\bk{\S_{\v m}}$ takes only a short
-time independent of system size.  Finally, the relation
-$\S_{\v m}^\dag=\S_{\v m^*}$ for $\v m^*\equiv\p{m_-,m_\z,m_+}$ can be
-exploited to cut both the number of matrix elements $T_{\v m\v n}$ and
-expectation values $\bk{\S_{\v m}}_{t=0}$ that we need to compute
-explicitly roughly in half.
-
-\bibliography{\jobname}
-
-\newpage
-\appendix
+\eqref{eq:Sz_OAT} are sufficient to compute the spin squeezing
+parameter at any time throughout evolution of the initial state
+$\ket{\X}$ under $H_{\t{OAT}}$.
 
@@
 
-The appendices in this text make ubiquitous use of various spin
+The appendices in this work make ubiquitous use of various spin
 operator identities; we collect and derive some basic identities here
-for reference.  Defining the individual spin-1/2 operators
-\begin{align}
-  s_{n=\x,\y,\z} \equiv \f12 \sigma_n,
-  &&
-  s_\pm \equiv s_\x \pm i s_\y,
-  \label{eq:single_spin}
-\end{align}
-in terms of Pauli operators $\sigma_n$, we denote the operator which
-acts with $s_n$ on a spin indexed by $j$ and trivially (i.e.~with the
-identity $\1$) on all other spins by $s_n^{(j)}$.  We then define the
-collective spin operators
-\begin{align}
-  S_n \equiv \sum_j s_n^{(j)}.
-  \label{eq:collective_spin}
-\end{align}
-Despite our working definition in \eqref{eq:collective_spin}, the
-identities we will derive for collective spin operators alone apply
-just as well to large-spin operators that cannot be expressed as the
-sum of individual spin-1/2 operators.
-
-The elementary commutation relations between spin
+for reference.  Note that despite the working definition of collective
+spin operators from $S_n=\sum_js_n^{(j)}$, the identities we will
+derive involving only collective spin operators apply just as well to
+large-spin operators that cannot be expressed as the sum of individual
+spin-1/2 operators.  The elementary commutation relations between spin
 operators are, with $\bmu\equiv-\mu\in\set{+1,-1}$ for brevity,
@@
   = \cdots
-  = ms_\mu^{(j)} S_\mu^{m-1} + S_\mu^m \mu s_\z^{(j)},
+  = S_\mu^m \mu s_\z^{(j)} + ms_\mu^{(j)} S_\mu^{m-1},
   \label{eq:push_z_mu_sS}
@@
 $\mu s_\z^{(j)}$ and $\mu S_\z$ rather than $s_\z^{(j)}$ and $S_\z$.
-Summing over the single-spin index $j$ in both of these cases gives us
-the purely collective-spin versions of these identities:
+Summing over the single-spin index $j$ in both of the cases above
+gives us the purely collective-spin versions of these identities:
 \begin{align}
@@
   D_{mn;\mu}^{(k)}
-  &\equiv D_{m-k,n;\mu}S_\mu^k \\
   &= \p{m-k}n S_\bmu^{n-1} S_\mu^{m-k-1}
@@
 
-Observing that we always have
-$0\le k_1\le k_2\le\cdots\le k_p\le m-p-1$ in \eqref{eq:E_mn_p}, we
-can rearrange the order of the sums and relabel $k_p\to\ell$ to get
+Observing that in \eqref{eq:E_mn_p} we always have
+$0\le k_1\le k_2\le\cdots\le k_p\le m-p-1$, we can rearrange the order
+of the sums and relabel $k_p\to\ell$ to get
 \begin{align}
@@
 
-\section{Changing bases for collective-spin operators}
-
-For a collective-spin system composed of many small spins, the
+\section{Changing bases for collective spin operators}
+
+For a collective spin system composed of many small spins, the
 operators $\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$ keep track
@@
 with $k<\abs{\v m}$.  To cut down on this overhead, we change our
-basis for collective-spin operators from $\S_{\v m}$ to the
+basis for collective spin operators from $\S_{\v m}$ to the
 purely-$\abs{\v m}$-body operators
@@
 are equal up to a permutation of indices if
-$\tilde{\v\rho_1}=\tilde{\v\rho_2}$.  The degeneracy of terms in
+$\tilde{\v\rho}_1=\tilde{\v\rho}_2$.  The degeneracy of terms in
 \eqref{eq:Q_K_rho}, i.e.~the number of $\v\rho$ which are consistent
@@
 
+TODO: comment on the use of this new basis vs.~the old one.
+
 \end{document}

### 2018-11-18 soc_squeezing/collective_spin.tex
fixed up some text, removed extra bib entry

--- 2018-11-16 soc_squeezing/cumulant_expansions.tex
@@
 \newcommand{\bk}{\Braket} % shorthand for braket notation
-\renewcommand{\d}{\text{d}} % "d" for integration measure
 
@@
 \newcommand{\bnu}{{\bar\nu}}
-\newcommand{\floor}[1]{\left\lfloor{#1}\right\rfloor}
 
@@
 
-\title{Short-time simulation of collective quantum spin systems with
-  decoherence}
+\title{Short-time simulation of open collective quantum spin systems}
 
@@
 simulations become dominated by incoherent jumps that generate large
-numbers of distinct trajectories over which to average.
+numbers of distinct trajectories over which to average.  [Unclear?]
 
@@
 \end{align}
-The commutator in Eq.~\eqref{eq:H_general} can be determined from the
+The commutator in Eq.~\eqref{eq:EOM} can be determined from the
 structure factors of $f_{\v n\v m\v\ell}$ of the general product
@@
 for some weight measure $w$, e.g.~$w\p{\v n}=\abs{\v n}$, and cutoff
-$W$.  The truncation in \eqref{eq:weight_truncation} will close the
-system of differential equations defined by \eqref{eq:EOM}, and allow
-us to solve it using standard numerical methods.  Initial conditions
-for this system of differential equations, i.e.~expectation values of
-collective spin operators with respect to states that are simple to
-prepare experimentally, are provided in Appendix
+$W$.  The truncation in Eq.~\eqref{eq:weight_truncation} will close
+the system of differential equations defined by Eq.~\eqref{eq:EOM},
+and allow us to solve it using standard numerical methods.  Initial
+conditions for this system of differential equations, i.e.~expectation
+values of collective spin operators with respect to states that are
+simple to prepare experimentally, are provided in Appendix
 \ref{sec:initial_conditions}.
@@
   = \sum_{k\ge0} \f{t^k}{k!}
-  \sum_{\v n} T_{\v m\v n;k} \bk{\S_{\v n}}_{t=0},
+  \sum_{\v n} T^k_{\v m\v n} \bk{\S_{\v n}}_{t=0},
   \label{eq:time_series}
 \end{align}
-where the matrix elements of the $k$-th time derivative are given by
-\begin{align}
-  T_{\v m\v n;0} \equiv \delta_{\v m\v n},
-  &&
-  T_{\v m\v n;1} \equiv T_{\v m\v n},
-  &&
-  T_{\v m\v n;k>1}
+where the matrix elements $T^k_{\v m\v n}$ of the $k$-th time
+derivative are
+\begin{align}
+  T^0_{\v m\v n} \equiv \delta_{\v m\v n},
+  &&
+  T^1_{\v m\v n} \equiv T_{\v m\v n},
+  &&
+  T^{k>1}_{\v m\v n}
   \equiv \sum_{\v p_1,\v p_2,\cdots,\v p_{k-1}}
@@
 \end{align}
-where $\delta_{\v m\v n}=1$ if $\v m=\v n$ and 0 otherwise.  For
+with $\delta_{\v m\v n}=1$ if $\v m=\v n$ and 0 otherwise.  For
 sufficiently short times, we can truncate the series in
@@
   \to \sum_{k=0}^M \f{t^k}{k!}
-  \sum_{\v n} T_{\v m\v n;k} \bk{\S_{\v n}}_{t=0}.
-\end{align}
-Unlike the truncation strategy in Eq.~\eqref{eq:weight_truncation},
-the nonzero entries $T_{\v m\v n;k}$ for $k=0,1,\cdots,M$ now
-explicitly tell us which operators are relevant for computing the
-expectation value $\bk{\S_{\v m}}$ to a fixed order $M$.  Furthermore,
-after computing all necessary matrix elements $T_{\v m\v n;k}$ and
-expectation values $\bk{\S_{\v n}}_{t=0}$, computing $\bk{\S_{\v m}}$
-takes only an $\O\p{M}$ amount of computer time independent of the
-system time $t$.  Finally, we note that the relation
+  \sum_{\v n} T^k_{\v m\v n} \bk{\S_{\v n}}_{t=0}.
+\end{align}
+Unlike the truncation in Eq.~\eqref{eq:weight_truncation}, the nonzero
+entries $T^k_{\v m\v n}$ for $k=0,1,\cdots,M$ now explicitly tell us
+which operators are relevant for computing the expectation value
+$\bk{\S_{\v m}}$ to a fixed order $M$.  Note that the relation
 $\S_{\v m}^\dag=\S_{\v m^*}$ for $\v m^*\equiv\p{m_-,m_\z,m_+}$ can be
@@
 \end{align}
-where $\Gamma$ is a collective spin operator with weight one, i.e.~a
+where $\Gamma$ is a collective spin operator with unit weight, i.e.~a
 linear combination of $S_\z$, $S_\mu$, and $S_\bmu$.
@@
 
-Initial conditions for all collective spin operators can be determined
-from the expansions
+Initial conditions for all collective spin operators with respect to
+(spin-polarized) Gaussian states in the Dicke manifold spanned by
+$\set{\ket{m}:S_\z\ket{m}=m\ket{m}}$ can be determined from the
+expansions
 \begin{align}
@@
 \end{align}
-with
-\begin{align}
-  g_\mu\p{S,k} \equiv \sqrt{\p{S-\mu k}\p{S+\mu k+1}},
-\end{align}
-and the initial states
+with $\mu\in\set{+1,-1}$ and
+\begin{align}
+  g_\mu\p{S,k} \equiv \sqrt{\p{S-\mu k}\p{S+\mu k+1}}.
+\end{align}
+Given the initial states
 \begin{align}
@@
 \end{align}
-using which we can compute, for $\mu,\nu\in\set{+1,-1}$,
-$\bmu\equiv-\mu$,
+where $S_\x\ket{\X}=S\ket{\X}$, in terms of $\nu\in\set{+1,-1}$ and
+$\bmu\equiv-\mu$ we then have
 \begin{align}
@@
 \begin{align}
-  \ket{\phi} \equiv e^{-i\phi S_\z}\ket{\X},
-  &&
-  \p{\cos\phi S_\x + \sin\phi S_\y} \ket\phi = S \ket\phi,
+  \ket{\tilde\phi} \equiv e^{-i\phi S_\z}\ket{\X},
+  &&
+  \p{\cos\phi S_\x + \sin\phi S_\y} \ket{\tilde\phi}
+  = S \ket{\tilde\phi},
 \end{align}
@@
 \begin{align}
-  \bk{\phi|S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n|\phi}
+  \bk{\tilde\phi|S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n|\tilde\phi}
   = e^{i\mu\p{\ell-n}\phi} \bk{\X|S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n|\X}.
@@
     + \f{\p{N-\ell}\p{N-n}}{N} \beta_{N\ell mn}},
-\end{align}
-where $\alpha$ and $\beta$ are defined in terms of the generalized
-hypergeometric function $\F$ by
+  \label{eq:X_hyper}
+\end{align}
+where $\alpha_{N\ell n}$ and $\beta_{N\ell mn}$ are defined in terms
+of the generalized hypergeometric function $\F$ by
 \begin{align}
@@
 $r_{\mu\nu}=\abs{\set{K^{\mu\nu}_a}}$ for brevity; and
-$\eta_{\mu\nu\rho}$ is defined by
+$\eta_{\mu\nu\rho}$ is a structure factor defined by
 $s_\mu s_\nu=\sum_\rho\eta_{\mu\nu\rho}s_\rho$.  Unlike the sums over
@@
 \end{align}
-In addition, the multiplication table $\eta_{\mu\nu\kappa}$ may
-generally contain zeros that we can preemptively avoid with
-restrictions on $\v r$ and $\v\rho$.  Recalling that $r_{\mu\nu}$
-counts the number of operators addressed by
+In addition, the structure factor $\eta_{\mu\nu\kappa}$ may generally
+contain zeros that we can preemptively avoid with restrictions on
+$\v r$ and $\v\rho$.  Recalling that $r_{\mu\nu}$ counts the number of
+operators addressed by
 $s_\mu s_\nu=\sum_\kappa\eta_{\mu\nu\kappa} s_\kappa$, if
@@
 all $\mu,\nu$ with $s_\mu s_\nu=0$.  Finally, we can avoid the
-nullifying effect of any additional zeros in the multiplication table
-$\eta_{\mu\nu\kappa}$, which appear as factors in
+nullifying effect of zeros from $\eta_{\mu\nu\kappa}$
 $\v\eta_{\v\rho}\propto\p{\eta_{\mu\nu\kappa}}^{\rho^{\mu\nu}_\kappa}$
-as defined in \eqref{eq:eta_rho}, by restricting
-$\rho^{\mu\nu}_\kappa=0$ for all $\mu,\nu,\kappa$ with
+by fixing $\rho^{\mu\nu}_\kappa=0$ for all $\mu,\nu,\kappa$ with
 $\eta_{\mu\nu\kappa}=0$.  These restrictions are sufficient to ensure

### 2018-11-19 soc_squeezing/collective_spin.tex
simplified some text and removed hypergeometric solutions for expectation values

@@
 simulations become dominated by incoherent jumps that generate large
-numbers of distinct trajectories over which to average.  [Unclear?]
+numbers of distinct trajectories over which to average.
 
 In this work, we develop a new method to simulate short-time dynamics
-of collective spin systems with arbitrary-strength
-permutationally-symmetric decoherence.  This method is based on a
-short-time expansion of exact solutions to the equations of motion for
-Heisenberg operators.  Evaluating this expansion requires use of the
-structure factors of a collective spin operator algebra; the
-calculation of these structure factors is one of the main technical
-results of this work.
+of collective spin systems with permutationally-symmetric decoherence.
+This method is based on a short-time expansion of exact solutions to
+the equations of motion for Heisenberg operators.  Evaluating this
+expansion requires use of the structure factors of a collective spin
+operator algebra; the calculation of these structure factors is one of
+the main technical results of this work.
 
@@
 distinct spin-1/2 particles.  Defining individual spin-1/2 operators
-$s_{n=\x,\y,\z}\equiv\sigma_n/2$ and $s_\pm\equiv s_\x\pm is_\y$ with
-Pauli operators $\sigma_n$, we denote an operator which acts with
-$s_n$ on the spin indexed by $j$ and trivially (i.e.~with the identity
-$\1$) on all other spins by $s_n^{(j)}$.  We then define the
-collective spin operators $S_n\equiv\sum_{j=1}^Ns_n^{(j)}$ for
+$s_{n=\x,\y,\z}\equiv\sigma_n/2$ and
+$s_\pm\equiv s_\x\pm is_\y=\sigma_\pm$ with Pauli operators
+$\sigma_n$, we denote an operator which acts with $s_n$ on the spin
+indexed by $j$ and trivially (i.e.~with the identity $\1$) on all
+other spins by $s_n^{(j)}$.  We then define the collective spin
+operators $S_n\equiv\sum_{j=1}^Ns_n^{(j)}$ for
 $n\in\set{\x,\y,\z,+,-}$.  Identifying the set $\set{\S_{\v m}}$ as a
@@
   \D\p{\J;\O}
-  \equiv \sum_{J\in\J} \p{J^\dag \O J - \f12\sp{J^\dag J,\O}_+}.
-\end{align}
-The commutator in Eq.~\eqref{eq:EOM} can be determined from the
-structure factors of $f_{\v n\v m\v\ell}$ of the general product
+  \equiv \sum_{J\in\J} \p{J^\dag \O J - \f12\sp{J^\dag J,\O}_+}
+\end{align}
+is the Heisenberg picture dissipator, or Lindblad superoperator, for
+all jump operators $J\in\J$.  Decoherence via uncorrelated decay of
+individual spins at a rate $\gamma_-$, for example, would be described
+by the set of jump operators
+$\J_-=\set{\sigma_-^{(j)}:j=1,2,\cdots,N}$ with
+$\gamma_{\J_-}=\gamma_-$.  The commutator in Eq.~\eqref{eq:EOM} can be
+determined from the structure factors of $f_{\v n\v m\v\ell}$ of the
+general product
 $\S_{\v n}\S_{\v m}=\sum_{\v\ell}f_{\v n\v m\v\ell}S_{\v\ell}$ worked
@@
 any method tracking only a subset of all relevant operators.  We
-therefore devise an alternate truncation strategy built around point
-\ref{pt:limitation}.
+therefore devise an alternate truncation strategy built around
+limitation \ref{pt:limitation}.
 
@@
 Unlike the truncation in Eq.~\eqref{eq:weight_truncation}, the nonzero
-entries $T^k_{\v m\v n}$ for $k=0,1,\cdots,M$ now explicitly tell us
+matrix elements $T^k_{\v m\v n}$ for $k=0,1,\cdots,M$ now tell us
 which operators are relevant for computing the expectation value
@@
   = \chi S_\z^2
-  = \f14 \chi \sum_{j,k} \sigma_\z^{(j)} \sigma_\z^{(k)}
   = \f12 \chi \sum_{j<k} \sigma_\z^{(j)} \sigma_\z^{(k)} + \f14 N \chi,
 \end{align}
-where $\sigma_z^{(j)}$ represents a Pauli-$z$ operator acting on spin
-$j$.  This Hamiltonian admits an exact solution previously worked out
-in Ref.~[\citenum{foss-feig2013nonequilibrium}].  Adapting the
-expectation values therein with respect to the initial state
-$\ket{\X}$ satisfying $S_\x\ket{\X}=S\ket{\X}$, in terms of
-$S\equiv N/2$ and $\mu,\nu\in\set{+1,-1}$ we find that
+where $\sigma_\z^{(j)}$ represents a Pauli-$z$ operator acting on spin
+$j$.  This Hamiltonian is a special case of the Ising Hamiltonian
+previously solved in Ref.~[\citenum{foss-feig2013nonequilibrium}] via
+exact, analytical treatment of the quantum jump Monte Carlo method for
+computing expectation values.  The solution therein accounts for
+coherent evolution in addition to decoherence via uncorrelated
+single-spin decay, excitation, and dephasing respectively at rates
+$\gamma_-$, $\gamma_+$, and $\gamma_\z$ (denoted by $\Gamma_{\t{ud}}$,
+$\Gamma_{\t{du}}$, and $\Gamma_{\t{el}}$ in
+Ref.~[\citenum{foss-feig2013nonequilibrium}]).  Letting $S\equiv N/2$
+and $\mu,\nu\in\set{+1,-1}$, we adapt expectation values computed in
+Ref.~[\citenum{foss-feig2013nonequilibrium}] for the initial state
+$\ket{\X}$ satisfying $S_\x\ket{\X}=S\ket{\X}$ evolving under
+$H_{\t{OAT}}$, finding
 \begin{align}
   \bk{S_+}
-  &= \sum_j\bk{\sigma_+^{(j)}}
-  = S e^{-\Gamma t} \Phi\p{\chi,t}^{N-1}, \label{eq:S+_OAT} \\
+  &= S e^{-\Gamma t} \Phi\p{\chi,t}^{N-1}, \label{eq:S+_OAT} \\
   \bk{S_\mu S_\z}
-  &= \f12\sum_j\bk{\sigma_\mu^{(j)}\sigma_\z^{(j)}}
-  + \f12\sum_{j\ne k} \bk{\sigma_\mu^{(j)} \sigma_\z^{(k)}} \\
   &= -\f{\mu}{2}\bk{S_\mu} + S \p{S-\f12} e^{-\Gamma t}
@@
   \bk{S_\mu S_\nu}
-  &= \sum_j \bk{\sigma_\mu^{(j)} \sigma_\nu^{(j)}}
-  + \sum_{j\ne k} \bk{\sigma_\mu^{(j)} \sigma_\nu^{(k)}} \\
   &= \delta_{\mu,-\nu} \p{S + \mu\bk{S_\z}}
@@
 \end{align}
-defined respectively in terms of uncorrelated single-spin decay,
-excitation, and dephasing rates $\gamma_-$, $\gamma_+$, $\gamma_\z$
-(respectively $\Gamma_{\t{ud}}$, $\Gamma_{\t{du}}$, and
-$\Gamma_{\t{el}}$ in Ref.~[\citenum{foss-feig2013nonequilibrium}]).
 The last ingredients we need to compute spin squeezing at any time are
@@
 The expectation values in \eqref{eq:S+_OAT}-\eqref{eq:SS+-_OAT} and
-\eqref{eq:Sz_OAT} are sufficient to compute the spin squeezing
-parameter at any time throughout evolution of the initial state
-$\ket{\X}$ under $H_{\t{OAT}}$.
+\eqref{eq:Sz_OAT} are sufficient to compute the spin squeezing at any
+time throughout evolution of the initial state $\ket{\X}$ under
+$H_{\t{OAT}}$.
 
@@
 \begin{align}
-  \p{\nu S_z}^\ell S_\mu^m
+  \p{\nu S_\z}^\ell S_\mu^m
   = S_\mu^m \p{\mu\nu m+\nu S_\z}^\ell,
   &&
-  S_\mu^m \p{\nu S_z}^\ell
+  S_\mu^m \p{\nu S_\z}^\ell
   = \p{-\mu\nu m+\nu S_\z}^\ell S_\mu^m.
@@
 \end{align}
-and
+1and
 \begin{align}
@@
   \f{k^m \p{S-\mu k}!}
-  {\p{S+\mu k}!\p{S-\mu k-\ell}!\p{S-\mu k-n}!} \\
+  {\p{S+\mu k}!\p{S-\ell-\mu k}!\p{S-n-\mu k}!} \\
   &= \f{N!}{2^N} \sum_{k=-S}^{S-\max\set{\ell,n}}
-  \f{k^m \p{S-k}!}{\p{S+k}!\p{S-k-\ell}!\p{S-k-n}!}.
+  \f{k^m \p{S-k}!}{\p{S+k}!\p{S-\ell-k}!\p{S-n-k}!} \\
+  &= \p{-1}^m\f{N!}{2^N} \sum_{k=0}^{N-\max\set{\ell,n}}
+  \f{\p{S-k}^m \p{N-k}!}{k!\p{N-\ell-k}!\p{N-n-k}!}.
 \end{align}
@@
 \end{align}
-Alternatively, with the help of Mathematica we can find
-\begin{align}
-  \bk{\X|S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n|\X}
-  = \f{\p{N!}^2}{2^N\p{N-\ell}!\p{N-n}!}
-  \sp{\p{-S}^m \alpha_{N\ell n}
-    + \f{\p{N-\ell}\p{N-n}}{N} \beta_{N\ell mn}},
-  \label{eq:X_hyper}
-\end{align}
-where $\alpha_{N\ell n}$ and $\beta_{N\ell mn}$ are defined in terms
-of the generalized hypergeometric function $\F$ by
-\begin{align}
-  \alpha_{N\ell n}
-  \equiv \F\p{\let\scriptstyle\textstyle
-    \substack{\ell-N,~n-N \\ -N}~; -1},
-\end{align}
-and
-\begin{align}
-  \beta_{N\ell mn}
-  \equiv \sum_{p=0}^{m-1} \p{-S}^p { m \choose p }
-  \tilde\F_{m-p-1}\p{\let\scriptstyle\textstyle
-    \substack{1+\ell-N,~1+n-N \\ 1-N}~; -1},
-\end{align}
-for
-\begin{align}
-  \tilde\F_q\p{\let\scriptstyle\textstyle
-    \substack{a,~b\\c}~; z}
-  \equiv \F\p{\let\scriptstyle\textstyle
-    \substack{2,2,\cdots,2,a,b \\ 1,1,\cdots,1,c}~; z}
-\end{align}
-with $q$ repetitions of 2 and 1.
 
@@
 For a collective spin system composed of many small spins, the
-operators $\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$ keep track
-of $\abs{\v m}$-body operators (for $\abs{\v m}\equiv\sum_\mu m_\mu$),
-but also carry lots of ``baggage'' in the form of $k$-body operators
-with $k<\abs{\v m}$.  To cut down on this overhead, we change our
-basis for collective spin operators from $\S_{\v m}$ to the
-purely-$\abs{\v m}$-body operators
+operators $\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$ contain
+genuinely $\abs{\v m}$-body operators (for
+$\abs{\v m}\equiv\sum_\mu m_\mu$), but also carry lots of ``baggage''
+in the form of $k$-body operators with $k<\abs{\v m}$.  To cut down on
+this overhead, we try to change our basis for collective spin
+operators from $\S_{\v m}$ to the purely-$\abs{\v m}$-body operators
 \begin{align}
@@
   P_{\v j} \equiv \prod_{\mu,a} s_\mu^{\p{j^\mu_a}},
-\end{align}
-where $\v m \equiv \p{m_\z,m_+,m_-}\in\mathbb{N}_0^3$ with $m_\mu$ the
+  \label{eq:S_P}
+\end{align}
+where $\v m \equiv \p{m_+,m_\z,m_-}\in\mathbb{N}_0^3$ with $m_\mu$ the
 number of $s_\mu$ operators in each term of $\S_{\v m}$; $\v j$ is an
@@
 Collecting terms in which $r_{\mu\nu}$ of the $s_\mu$ operators in
-$P_{\v j}$ address the same spin as a $s_\nu$ operator in $P_{\v k}$,
+$P_{\v j}$ address the same spin as an $s_\nu$ operator in $P_{\v k}$,
 we have
@@
   \sp{\prod_{\mu,\nu} r_{\mu\nu}!} \\
-  &= \sp{\prod_\mu \p{m_\mu}_{\sum_\rho r_{\mu\rho}}
-    \p{n_\mu}_{\sum_\rho r_{\rho\mu}}}
-  \sp{\prod_{\mu,\nu} r_{\mu\nu}!}^{-1},
-\end{align}
-with $\p{n}_k\equiv\prod_{j=0}^{k-1}\p{n-j}$ denoting a falling
-factorial; $\v J$ and $\v K$ are respectively ordered lists of indices
-for spins addressed by one and two single-spin operators in a term of
-the product $P_{\v j} P_{\v k}$ with fixed $\v r$; the sum over
+  &= \sp{\prod_\mu \f{m_\mu!}{\p{m_\mu-\sum_\rho r_{\mu\rho}}!}
+    \f{n_\mu!}{\p{n_\mu-\sum_\rho r_{\rho\mu}}!}}
+  \sp{\prod_{\mu,\nu} r_{\mu\nu}!}^{-1};
+\end{align}
+$\v J$ and $\v K$ are respectively ordered lists of indices for spins
+addressed by one and two single-spin operators in a term of the
+product $P_{\v j} P_{\v k}$ with fixed $\v r$; the sum over
 $\p{\v J,\v K;\v m,\v n,\v r}$ denotes a sum over all values of
@@
   &&
-  \set{J_a^\lambda} \cap \set{K_b^{\mu\nu}} = \set{};
-\end{align}
-and finally, similarly to $P_{\v j}$ in \eqref{eq:SS_PP} we define
+  \set{J_a^\lambda} \cap \set{K_b^{\mu\nu}} = \varnothing,
+\end{align}
+with $\varnothing$ denoting the empty set; and finally, similarly to
+$P_{\v j}$ in \eqref{eq:SS_P} we define
 \begin{align}
@@
   c_{\v\ell\v\rho}
-  \equiv \p{ N - \sum_\kappa\ell_\kappa }_{\sum_{\mu,\nu}\rho^{\mu\nu}_0}
+  \equiv \f{\p{N - \abs{\v\ell}}!}
+  {\p{N - \abs{\v\ell} - \sum_{\mu,\nu}\rho^{\mu\nu}_0}!}
 \end{align}

### 2018-11-25 soc_squeezing/collective_spin.tex
made intro text less specific to squeezing/OAT slightly simplified expectation value expressions

@@
-\documentclass[aps,11pt,notitlepage,nofootinbib]{revtex4-1}
+\documentclass[aps,11pt,notitlepage,nofootinbib,longbibliography]{revtex4-1}
 
@@
 window into many-body quantum phenomena that are both experimentally
-accessible\cite{fernholz2008spin, takano2009spin,
+accessible\cite{takano2009spin, chen2010heralded,
   riedel2010atomchipbased, gross2010nonlinear, martin2013quantum,
-  bohnet2016quantum, hosten2016quantum, norcia2018cavitymediated} and
-metrologically useful\cite{wineland1992spin, kitagawa1993squeezed,
-  ma2011quantum}.  Theoretical interest in such systems dates back to
-the mid-twentieth century with the introduction of the
-Lipkin-Meshkov-Glick (LMG) model as a toy model for testing many-body
-approximation methods in contemporary nuclear
-physics\cite{lipkin1965validity, meshkov1965validity,
+  bohnet2016quantum, hosten2016quantum, norcia2018cavitymediated,
+  jones2009magnetic} and metrologically
+useful\cite{riedel2010atomchipbased, gross2010nonlinear,
+  hosten2016quantum, jones2009magnetic, zhong2010simplified,
+  huang2015quantum, agarwal1997atomic, lau2014proposal,
+  wineland1992spin, kitagawa1993squeezed, ma2011quantum}.  Theoretical
+interest in such systems dates back to the mid-twentieth century with
+the introduction of the Lipkin-Meshkov-Glick (LMG) model as a toy
+model for testing many-body approximation methods in contemporary
+nuclear physics\cite{lipkin1965validity, meshkov1965validity,
   glick1965validity}.  Since then, collective system systems have
@@
   bhattacherjee2014nonequilibrium, klinder2015dynamical,
-  zhiqiang2017nonequilibrium, smale2018observation}, and most notably
-spin squeezing\cite{wineland1992spin, kitagawa1993squeezed,
-  ma2011quantum}.  With the development of advanced atomic trapping,
-cooling, and control techniques, collective spin systems have been
-experimentally realized in atomic ensembles\cite{fernholz2008spin,
-  takano2009spin}, Bose-Einstein
+  zhiqiang2017nonequilibrium, smale2018observation}, and precision
+metrology\cite{wineland1992spin, kitagawa1993squeezed, ma2011quantum,
+  zhong2010simplified}.  With the development of advanced atomic
+trapping, cooling, and control techniques, collective spin systems
+have been experimentally realized in atomic
+ensembles\cite{takano2009spin, chen2010heralded}, Bose-Einstein
 condensates\cite{riedel2010atomchipbased, gross2010nonlinear},
 ultracold Fermi gasses\cite{martin2013quantum}, trapped
-ions\cite{bohnet2016quantum}, and optical
-cavities\cite{hosten2016quantum, norcia2018cavitymediated}.
-
-One of the primary current motivations for studying collective spin
-systems is their capability to dynamically generate spin-squeezed
-states, which enable measurement of phase angles $\phi$ with an error
-$\Delta\phi$ below the standard quantum limit
-$\Delta\phi\sim1/\sqrt{N}$ set by quantum projection noise from the
-measurement of $N$ independent particles\cite{ma2011quantum}.  The
-nonlinear dynamics necessary to generate spin-squeezed states can be
-implemented with long-range collisional, photon-, or phonon-mediated
-interactions, typically resulting in a one-axis twising (OAT) model
-described by the Hamiltonian\cite{kitagawa1993squeezed}
-\begin{align}
-  H_{\t{OAT}} = \chi S_\z^2,
-  \label{eq:OAT}
-\end{align}
-where $S_\z\equiv\sum_{j=1}^N\sigma_\z^{(j)}/2$ is a collective
-spin-$z$ operator defined in terms of Pauli-$z$ operators
-$\sigma_\z^{(j)}$ addressing spin $j$.  The OAT model prepares a
-spin-squeezed state that allows measurement of a phase angle $\phi$
-with error $\Delta\phi\sim N^{-2/3}$ at times
-$t\sim\chi^{-1}N^{-2/3}$\cite{ma2011quantum}.  Although a truly
-collective spin model requires uniform, all-to-all interactions, as
-long as measurements do not distinguish between constituent particles
-even non-uniform systems can be effectively described by a uniform
-model with renormalized parameters\cite{hu2015entangled}.  Various
-works have proposed improvements over OAT through the addition of
-auxiliary control fields\cite{law2001coherent, liu2011spin,
-  huang2015twoaxis, muessel2015twistandturn} or a modified
-vacuum\cite{hu2017vacuum}, thereby realizing different collective spin
-models.
-
-While the OAT model admits exact
-solutions\cite{foss-feig2013nonequilibrium}, the same is not true of
-more general, non-Ising Hamiltonians.  Nonetheless, in the absence of
-decoherence, permutation symmetry and total spin conservation divide
-the total Hilbert space of a collective spin system into superseletion
-sectors that grow only linearly with system size, thereby admitting
-efficient classical simulation of its dynamics.  If decoherence is
-present, but sufficiently weak, dynamics of collective spin systems
-are still numerically solvable via ``quantum jump'' Monte Carlo
-methods\cite{plenio1998quantumjump, zhang2018montecarlo} (also known
-as ``quantum trajectory'' or ``Monte Carlo wavefunction'' methods)
-that can reproduce all expectation values of interest.  When
-decoherence is strong, however, or when the number of jump operators
-grows extensively with system size (e.g.~for single-spin decay), these
-Monte Carlo methods can take a prohibitively long time to converge, as
-simulations become dominated by incoherent jumps that generate large
-numbers of distinct trajectories over which to average.
+ions\cite{bohnet2016quantum}, optical cavities\cite{hosten2016quantum,
+  norcia2018cavitymediated}, and molecular
+sensors\cite{jones2009magnetic}.
+
+One of the primary motivations for studying collective spin systems is
+their application to quantum-enhanced metrology.  Quantum projection
+noise limits the error $\Delta\phi$ in the measurement of a phase
+angle $\phi$ with $N$ independent spins to
+$\Delta\phi\sim1/\sqrt{N}$\cite{wineland1992spin, itano1993quantum,
+  ma2011quantum}.  Collective spin systems enable a means to break
+through this limit via the preparation of many-body entangled states
+such as NOON\cite{jones2009magnetic, chen2010heralded,
+  zhong2010simplified}, spin-cat\cite{huang2015quantum,
+  agarwal1997atomic, lau2014proposal}, and spin-squeezed
+states\cite{takano2009spin, riedel2010atomchipbased,
+  gross2010nonlinear, hosten2016quantum, norcia2018cavitymediated,
+  wineland1992spin, kitagawa1993squeezed, ma2011quantum} that allow
+for measurement errors $\Delta\phi\sim1/N^\epsilon$ with
+$\epsilon>1/2$.  Generating such states requires nonlinear dynamics
+that are typically realized using long-range collisional, photon-, or
+phonon-mediated interactions.
+% typically resulting in a one-axis twising (OAT) model described by
+% the Hamiltonian\cite{kitagawa1993squeezed}
+% \begin{align}
+%   H_{\t{OAT}} = \chi S_\z^2,
+%   \label{eq:OAT}
+% \end{align}
+% where $S_\z\equiv\sum_{j=1}^N\sigma_\z^{(j)}/2$ is a collective
+% spin-$z$ operator defined in terms of Pauli-$z$ operators
+% $\sigma_\z^{(j)}$ addressing spin $j$.  The OAT model prepares a
+% spin-squeezed state that allows measurement of a phase angle $\phi$
+% with error $\Delta\phi\sim N^{-2/3}$ at times
+% $t\sim\chi^{-1}N^{-2/3}$\cite{ma2011quantum}.
+Although a truly collective spin model requires uniform, all-to-all
+interactions, as long as measurements do not distinguish between
+constituent particles even non-uniform systems can be effectively
+described by a uniform model with renormalized
+parameters\cite{hu2015entangled}.
+% Various works have proposed improvements over OAT through the
+% addition of auxiliary control fields\cite{law2001coherent,
+% liu2011spin, huang2015twoaxis, muessel2015twistandturn} or a
+% modified vacuum\cite{hu2017vacuum}, thereby realizing different
+% collective spin models.
+
+% While the OAT model admits exact
+% solutions\cite{foss-feig2013nonequilibrium}, the same is not true of
+% more general, non-Ising Hamiltonians.  Nonetheless,
+In the absence of decoherence, permutation symmetry and total spin
+conservation divide the total Hilbert space of a collective spin
+system into superseletion sectors that grow only linearly with system
+size $N$, thereby admitting efficient classical simulation of its
+dynamics.  Decoherence will generally violate total spin conservation
+and require the use of density operators, increasing the dimension of
+accessible state space to $\sim N^3$\cite{hartmann2016generalized,
+  xu2013simulating} and allowing for exact simulations of
+$N\lesssim100$ particles.  If decoherence is sufficiently weak,
+dynamics are numerically solvable for $N\lesssim10^4$ particles via
+``quantum jump'' Monte Carlo methods\cite{plenio1998quantumjump,
+  zhang2018montecarlo} (also known as ``quantum trajectory'' or
+``Monte Carlo wavefunction'' methods) that can reproduce all
+expectation values of interest.  When decoherence is strong, however,
+or when the number of jump operators grows extensively with system
+size (e.g.~for single-spin decay), these Monte Carlo methods can take
+a prohibitively long time to converge, as simulations become dominated
+by incoherent jumps that generate large numbers of distinct
+trajectories over which to average.
 
@@
 operator algebra; the calculation of these structure factors is one of
-the main technical results of this work.
+the main technical results of this work, which we hope will enable
+future studies to develop additional analytical insights into the
+structure and dynamics of collective spin systems.
 
@@
 for some weight measure $w$, e.g.~$w\p{\v n}=\abs{\v n}$, and cutoff
-$W$.  The truncation in Eq.~\eqref{eq:weight_truncation} will close
-the system of differential equations defined by Eq.~\eqref{eq:EOM},
-and allow us to solve it using standard numerical methods.  Initial
+$W$.  The truncation in Eq.~\eqref{eq:weight_truncation} closes the
+system of differential equations defined by Eq.~\eqref{eq:EOM}, and
+allow us to solve it using standard numerical methods.  Initial
 conditions for this system of differential equations, i.e.~expectation
@@
   the initial value of operators $\S_{\v m}$ with weight
-  $w\p{\v m}\ge W$ have a non-negligible contribution to operators of
-  interest.\label{pt:limitation}
+  $w\p{\v m}\ge W$ have a non-negligible contribution to expectation
+  values of interest.\label{pt:limitation}
 \end{enumerate*}
@@
 
-\section{Benchmarking}
+\section{Spin squeezing}
 
@@
 \end{align}
-1and
+and
 \begin{align}
   \bk{\X|S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n|\X}
-  &=  \f{\mu^m}{2^N} \sum_{k=-S+\delta_{\mu,-1}\max\set{\ell,n}}
-  ^{S-\delta_{\mu,1}\max\set{\ell,n}} k^m
+  &=  \f1{2^N} \sum_{k=-S+\delta_{\mu,-1}\max\set{\ell,n}}
+  ^{S-\delta_{\mu,1}\max\set{\ell,n}} \p{\mu k}^m
   \sp{{ N \choose S+k+\mu\ell } { N \choose S+k+\mu n }}^{1/2}
@@
   \prod_{q=0}^{n-1} g_\bmu\p{S,k+\mu\sp{n-q}} \\
-  &= \mu^m \f{N!}{2^N} \sum_{k=-S+\delta_{\mu,-1}\max\set{\ell,n}}
-  ^{S-\delta_{\mu,1}\max\set{\ell,n}}
-  \f{k^m \p{S-\mu k}!}
+  &= \f{N!}{2^N} \sum_{\mu k=-\mu S-\delta_{\mu,-1}\max\set{\ell,n}}
+  ^{\mu S-\delta_{\mu,1}\max\set{\ell,n}}
+  \f{\p{\mu k}^m \p{S-\mu k}!}
   {\p{S+\mu k}!\p{S-\ell-\mu k}!\p{S-n-\mu k}!} \\
@@
 \end{align}
-and say that
+and compute
 \begin{align}
@@
 with $\varnothing$ denoting the empty set; and finally, similarly to
-$P_{\v j}$ in \eqref{eq:SS_P} we define
+$P_{\v j}$ in \eqref{eq:S_P} we define
 \begin{align}

### 2018-11-28 soc_squeezing/paper/collective_spin.tex
added text to spin squeezing section: desription of numerical methods and benchmarking figures

--- 2018-11-25 soc_squeezing/collective_spin.tex
@@
-\documentclass[aps,11pt,notitlepage,nofootinbib,longbibliography]{revtex4-1}
+\documentclass[pra,reprint,longbibliography]{revtex4-1}
 
@@
 
+\newcommand{\up}{\uparrow}
+\newcommand{\dn}{\downarrow}
+
 \usepackage[inline]{enumitem} % for inline enumeration
@@
 \usepackage{graphicx} % for figures
-\usepackage{grffile} % help latex properly identify figure extensions
-\usepackage[caption=false]{subfig} % subfigures (via \subfloat[]{})
 \graphicspath{{./figures/}} % set path for all figures
@@
 dynamics are numerically solvable for $N\lesssim10^4$ particles via
-``quantum jump'' Monte Carlo methods\cite{plenio1998quantumjump,
-  zhang2018montecarlo} (also known as ``quantum trajectory'' or
-``Monte Carlo wavefunction'' methods) that can reproduce all
-expectation values of interest.  When decoherence is strong, however,
-or when the number of jump operators grows extensively with system
-size (e.g.~for single-spin decay), these Monte Carlo methods can take
-a prohibitively long time to converge, as simulations become dominated
-by incoherent jumps that generate large numbers of distinct
-trajectories over which to average.
+``quantum trajectory'' Monte Carlo methods\cite{plenio1998quantumjump,
+  zhang2018montecarlo} (also known as ``quantum jump'' or ``Monte
+Carlo wavefunction'' methods) that can reproduce all expectation
+values of interest.  When decoherence is strong, however, or when the
+number of jump operators grows extensively with system size (e.g.~for
+single-spin decay), these Monte Carlo methods can take a prohibitively
+long time to converge, as simulations become dominated by incoherent
+jumps that generate large numbers of distinct trajectories over which
+to average.
 
@@
 the main technical results of this work, which we hope will enable
-future studies to develop additional analytical insights into the
-structure and dynamics of collective spin systems.
+future studies to develop new analytical insights into the structure
+and dynamics of collective spin systems.
+
+[Paragraph or so about benchmarking and the discussion section?]
 
@@
 \end{align}
-where Hermiticity requires that $h_{m_+,m_\z,m_-}=h_{m_-,m_\z,m_+}^*$.
-The time evolution of a particular Heisenberg operator $\S_{\v m}$
-under a Hamiltonian of the form in Eq.~\eqref{eq:H_general} is then
-given by
+where Hermiticity requires that $h_{\v m}^*=h_{\v m^*}$ for
+$\v m^*\equiv\p{m_-,m_\z,m_+}$.  The time evolution of a particular
+Heisenberg operator $\S_{\v m}$ under a Hamiltonian of the form in
+Eq.~\eqref{eq:H_general} is then given by
 \begin{align}
   \f{d}{dt}\S_{\v m}
-  = i\sum_{\v n}h_{\v n}\sp{\S_{\v n},\S_{\v m}}_-
-  + \sum_\J \gamma_\J \D\p{\J;\S_{\v m}}
-  \equiv \sum_{\v n} T_{\v m\v n} \S_{\v n},
+  &= i\sum_{\v n}h_{\v n}\sp{\S_{\v n},\S_{\v m}}_-
+  + \sum_\J \gamma_\J \D\p{\J;\S_{\v m}} \\
+  &= \sum_{\v n} T_{\v m\v n} \S_{\v n},
   \label{eq:EOM}
 \end{align}
-where $\sp{X,Y}_\pm\equiv XY\pm YX$, $\J$ is a set of jump operators
-with a corresponding decoherence rate $\gamma_\J$, and
+where $\sp{X,Y}_\pm\equiv XY\pm YX$; $\J$ is a set of jump operators
+with a corresponding decoherence rate $\gamma_\J$;
 \begin{align}
@@
 is the Heisenberg picture dissipator, or Lindblad superoperator, for
-all jump operators $J\in\J$.  Decoherence via uncorrelated decay of
+all jump operators $J\in\J$; and $T_{\v m\v n}$ is a matrix element of
+the time derivative operator.  Decoherence via uncorrelated decay of
 individual spins at a rate $\gamma_-$, for example, would be described
@@
 $\S_{\v n}\S_{\v m}=\sum_{\v\ell}f_{\v n\v m\v\ell}S_{\v\ell}$ worked
-out in Appendix \ref{sec:prod_general}, while the effects of
-decoherence from jump operators (i.e.~elements of $\J$) of the form
-$\gamma^{(j)}=\sum_n\gamma_ns_n^{(j)}$ and $\Gamma=\sum_n\Gamma_nS_n$
-are respectively worked out in Appendices \ref{sec:decoherence_single}
-and \ref{sec:decoherence_collective}.  These ingredients are
-sufficient to compute matrix elements $T_{\v m\v n}$ of the time
-derivative operator $d/dt$ in Eq.~\eqref{eq:EOM} in most cases of
-interest.
+out in Appendices \ref{sec:identities}--\ref{sec:prod_general}, while
+the effects of decoherence from jump operators (i.e.~elements of $\J$)
+of the form $\gamma^{(j)}=\sum_n\gamma_ns_n^{(j)}$ and
+$\Gamma=\sum_n\Gamma_nS_n$ are respectively worked out in Appendices
+\ref{sec:sandwich_single}--\ref{sec:decoherence_collective}.  These
+ingredients are sufficient to compute matrix elements $T_{\v m\v n}$
+of the time derivative operator $d/dt$ in Eq.~\eqref{eq:EOM} in most
+cases of experimental relevance.
 
@@
 \begin{align}
-  T^0_{\v m\v n} \equiv \delta_{\v m\v n},
-  &&
-  T^1_{\v m\v n} \equiv T_{\v m\v n},
-  &&
+  T^0_{\v m\v n} &\equiv \delta_{\v m\v n}, \\
+  T^1_{\v m\v n} &\equiv T_{\v m\v n}, \\
   T^{k>1}_{\v m\v n}
-  \equiv \sum_{\v p_1,\v p_2,\cdots,\v p_{k-1}}
+  &\equiv \sum_{\v p_1,\v p_2,\cdots,\v p_{k-1}}
   T_{\v m\v p_1} T_{\v p_1\v p_2} T_{\v p_2\v p_3}
@@
   \sum_{\v n} T^k_{\v m\v n} \bk{\S_{\v n}}_{t=0}.
+  \label{eq:order_truncation}
 \end{align}
@@
 $\bk{\S_{\v m}}$ to a fixed order $M$.  Note that the relation
-$\S_{\v m}^\dag=\S_{\v m^*}$ for $\v m^*\equiv\p{m_-,m_\z,m_+}$ can be
-exploited to cut both the number of matrix elements $T_{\v m\v n}$ and
-expectation values $\bk{\S_{\v m}}_{t=0}$ that we need to explicitly
-compute roughly in half.
+$\S_{\v m^*}=\S_{\v m}^\dag$, which by Hermitian conjugation of
+Eq.~\eqref{eq:EOM} also implies that
+$T_{\v m^*\v n^*}=T_{\v m\v n}^*$, can be exploited to cut both the
+number of expectation values $\bk{\S_{\v m}}_{t=0}$ and matrix
+elements $T_{\v m\v n}$ that we need to explicitly compute roughly in
+half.
+
+In principle, the factorial suppression of terms at higher orders
+(i.e.~in $k$) of the expansion in Eq.~\eqref{eq:order_truncation}
+implies for any given time $t$, there exists a truncation order $M_t$
+for which truncation error is negligibly small.  In practice, only a
+fixed maximal truncation order $\overline M$ is accessible with
+limited computational resources, such that
+Eq.~\eqref{eq:order_truncation} only allows computation of the
+correlator $\bk{\S_{\v m}}$ to a maximal time $t_{\overline M}$.  The
+maximal accessible truncation order $\overline M$ generally depends on
+the choice of operator basis for collective spin systems, which we
+discuss briefly at the end of Appendix \ref{sec:new_basis}.  As we
+will see in the following sections, our choice of the basis
+$\set{\S_{\v m}}$ allows us to accurately compute spin correlators to
+times that are relevant e.g.~for spin squeezing protocols.
 
@@
 
-Our measure of spin squeezing is
-\begin{align}
-  \xi^2 \equiv \f{N}{\abs{\bk{\v S}}^2}
+To benchmark our method for computing collective spin correlators, we
+consider three collective spin models known to dynamically generate
+spin-squeezed states: the one-axis twisting (OAT), two-axis twisting
+(TAT), and twist-and-turn (TNT) models described by the collective
+spin Hamiltonians\cite{kitagawa1993squeezed, micheli2003manyparticle}
+\begin{align}
+  \label{eq:OAT}
+  H_{\t{OAT}} &= \chi S_\z^2, \\
+  \label{eq:TAT}
+  H_{\t{TAT}} &= \f{\chi}{3} \p{S_\z^2 - S_\y^2}, \\
+  \label{eq:TNT}
+  H_{\t{TNT}} &= \chi S_\z^2 + \Omega S_\x,
+\end{align}
+where we include a factor of $1/3$ in the TAT Hamiltonian because it
+naturally appears in realistic proposals to experimentally realize
+TAT\cite{liu2011spin, huang2015twoaxis}.  For simplicity, we further
+fix $\Omega=S\chi$ (with $S\equiv N/2$ throughout this work) to the
+critical value known to maximize the entanglement generation rate of
+of TNT in the large-$N$ limit\cite{micheli2003manyparticle,
+  sorelli2015fast}.  Note that the OAT model is a special case of the
+zero-field Ising model, whose quantum dynamics in the presence of
+decoherence via uncorrelated decay, excitation, and dephasing of
+individual spins have been previously solved
+analytically\cite{foss-feig2013nonequilibrium} (see Appendix
+\ref{sec:OAT}, as well as the Supplementary Material of
+Ref.~[\citenum{bohnet2016quantum}] for more general multi-spin
+correlators).  We will benchmark against these analytical OAT results
+wherever applicable.
+
+The Hamiltonians in Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT} squeeze the
+initial product state $\ket{\X}\propto\p{\ket\up+\ket\dn}^{\otimes N}$
+with $S_\x\ket{\X}=S\ket{\X}$.  Our measure of spin squeezing is the
+directionally-unbiased squeezing parameter determined by the maximal
+gain in resolution $\Delta\phi$ of a phase angle $\phi$ over that
+achieved by any spin-polarized product state
+(e.g.~$\ket{\X}$)\cite{wineland1992spin, ma2011quantum},
+\begin{align}
+  \xi^2
+  \equiv \f{\p{\Delta\phi_{\t{min}}}^2}{\p{\Delta\phi_{\t{polarized}}}^2}
+  = \f{N}{\abs{\bk{\v S}}^2}
   \min_{\uv n\perp\bk{\v S}} \bk{\p{\v S\c\uv n}^2},
@@
 \end{align}
-where $N$ is the total number of spins, $\v S$ is the collective spin
-vector, and the minimization is performed over all unit vectors
-$\uv n$ in the plane orthogonal to the mean spin vector $\bk{\v S}$.
-This squeezing parameter is entirely determined by one- and two-spin
-correlators of the form $\bk{S_\alpha}$ and $\bk{S_\alpha S_\beta}$.
-
-\section{Discussion}
+where $\v S\equiv\p{S_\z,S_\x,S_\y}$ is a collective spin vector, and
+the minimization is performed over all unit vectors $\uv n$ in the
+plane orthogonal to the mean spin vector $\bk{\v S}$.  This squeezing
+parameter is entirely determined by one- and two-spin correlators of
+the form $\bk{S_m}$ and $\bk{S_m S_n}$.  In the case of coherent
+dynamics under TAT or TNT, these correlators are obtainable via exact
+simulations of quantum dynamics in the $\p{N+1}$-dimensional Dicke
+manifold of states $\set{\ket{S,m}}$ with net spin $S$ and spin
+projection $m$ onto the $z$ axis, i.e.~with
+$\bk{S,m|\v S^2|S,m}=S\p{S+1}$ and $\bk{S,m|S_\z|S,m}=m$ for
+$\abs{m}\le S$.  In the presence of permutationally-symmetric
+decoherence, these correlators are obtainable with the collective-spin
+quantum trajectory Monte Carlo method recently developed in
+ref.~\cite{zhang2018montecarlo}.  These exact and quantum trajectory
+simulations will serve as our methods for benchmarking the expansion
+in Eq.~\eqref{eq:order_truncation}.
+
+\begin{figure}
+  \centering
+  \includegraphics{coherent.pdf}
+  \caption{Spin squeezing of $N=10^4$ spins under coherent evolution,
+    as computed using exact methods (solid lines) and the expansion in
+    Eq.~\eqref{eq:order_truncation} with $M=40$ (dashed lines).  Solid
+    circles mark the times for TAT and TNT after which the latter
+    method gives an unphysical result with $\xi^2<0$.  Both methods
+    agree exactly in the OAT case for all times shown.}
+  \label{fig:coherent}
+\end{figure}
+
+Figure \ref{fig:coherent} compares the squeezing parameter $\xi^2$ for
+$N=10^4$ spins initially in the state $\ket{\X}$ coherently evolved
+under the Hamiltonians in Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT}, as
+computed via both benchmarking simulations and the prescription in
+Eq.~\eqref{eq:order_truncation} with $M=40$.  Similarly, Figure
+\ref{fig:weak_decoherence} compares squeezing in the same scenario
+with the addition of decoherence through uncorrelated emission,
+absorption, and dephasing of individual spins at a rate $\chi$,
+respectively described by the sets of jump operators
+$\J_n=\set{s_n^{(j)}}$ with corresponding decoherence rates
+$\gamma_n=\chi$ for $n\in\set{-,+,\z}$.  The squeezing results shown
+in Figures \ref{fig:coherent} and \ref{fig:weak_decoherence} were
+computed in a rotated basis with $\p{s_\z,s_\x}\to\p{s_\x,-s_\z}$ and
+$\ket\X\to\ket{-\Z}\equiv\ket\dn^{\otimes N}$; the only effect of this
+rotation on the results presented in these figures is to roughly
+double the time span of near-perfect agreement in squeezing via TNT.
+The reason for different results in rotated bases has to do with the
+breakdown of the expansion in Eq.~\eqref{eq:order_truncation}, which
+we discuss further in Section \ref{sec:breakdown}.
+
+[Comment on near-perfect agreement right up until a sudden and drastic
+departure from the correct result.  Reference figure showing squeezing
+for large system with strong decoherence.  Mention that this method
+was used to put a lower bound on squeezing via TAT in our
+SOC+interactions squeezing paper.]
+
+\section{Many-body correlators}
 
 Many-body correlators and cumulants \`a la
-Ref.~\cite{schweigler2017solving}?
-
-\section{Conclusion}
-
-\bibliography{\jobname}
+Ref.~[\citenum{schweigler2017solving}]?
+
+\section{Breakdown}
+\label{sec:breakdown}
+
+\section{Conclusions}
 
 \newpage
+\onecolumngrid
 \appendix
@@
 where $\sigma_\z^{(j)}$ represents a Pauli-$z$ operator acting on spin
-$j$.  This Hamiltonian is a special case of the Ising Hamiltonian
+$j$.  This model is a special case of the zero-field Ising Hamiltonian
 previously solved in Ref.~[\citenum{foss-feig2013nonequilibrium}] via
-exact, analytical treatment of the quantum jump Monte Carlo method for
-computing expectation values.  The solution therein accounts for
-coherent evolution in addition to decoherence via uncorrelated
-single-spin decay, excitation, and dephasing respectively at rates
-$\gamma_-$, $\gamma_+$, and $\gamma_\z$ (denoted by $\Gamma_{\t{ud}}$,
-$\Gamma_{\t{du}}$, and $\Gamma_{\t{el}}$ in
+exact, analytical treatment of the quantum trajectory Monte Carlo
+method for computing expectation values.  The solution therein
+accounts for coherent evolution in addition to decoherence via
+uncorrelated single-spin decay, excitation, and dephasing respectively
+at rates $\gamma_-$, $\gamma_+$, and $\gamma_\z$ (denoted by
+$\Gamma_{\t{ud}}$, $\Gamma_{\t{du}}$, and $\Gamma_{\t{el}}$ in
 Ref.~[\citenum{foss-feig2013nonequilibrium}]).  Letting $S\equiv N/2$
@@
 \end{align}
-The last ingredients we need to compute spin squeezing at any time are
-$\bk{S_\z}$ and $\bk{S_\z^2}$.  As these operators commute with the
-OAT Hamiltonian, their evolution is governed entirely by decoherence
-(see Appendix \ref{sec:decay_single}), which means
+In order to compute spin squeezing as measured by $\xi^2$ in
+\eqref{eq:squeezing}, we additionally need analytical expressions for
+$\bk{S_\z}$ and $\bk{S_\z^2}$.  As these operators commute with both
+the OAT Hamiltonian and the single-spin operators $\sigma_\z^{(j)}$,
+their evolution is governed entirely by decay-type decoherence (see
+Appendix \ref{sec:decay_single}), which means
 \begin{align}
@@
   \bk{S_\z}
-  = S\p{\f{\gamma_+-\gamma_-}{\gamma_++\gamma_-}}
-  \p{1-e^{-\p{\gamma_+ + \gamma_-} t}},
-  &&
-  \bk{S_\z^2} = S \sp{\f12 + \p{S-\f12} \f{\bk{S_\z}^2}{S^2}}.
+  &= S\p{\f{\gamma_+-\gamma_-}{\gamma_++\gamma_-}}
+  \p{1-e^{-\p{\gamma_+ + \gamma_-} t}}, \\
+  \bk{S_\z^2} &= S \sp{\f12 + \p{S-\f12} \f{\bk{S_\z}^2}{S^2}}.
   \label{eq:Sz_OAT}
@@
 The expectation values in \eqref{eq:S+_OAT}-\eqref{eq:SS+-_OAT} and
-\eqref{eq:Sz_OAT} are sufficient to compute the spin squeezing at any
-time throughout evolution of the initial state $\ket{\X}$ under
-$H_{\t{OAT}}$.
+\eqref{eq:Sz_OAT} are sufficient to compute the spin squeezing
+parameter $\xi^2$ in \eqref{eq:squeezing} at any time throughout
+evolution of the initial state $\ket{\X}$ under $H_{\t{OAT}}$, under
+appropriate assumptions about the relevant sources of decoherence.
 
@@
 all $s_\z^{(j)}$ operators on the right, as such an ordering will
-prove convenient for the calculations in this section.  This choice of
-ordering compels us to expand
+prove convenient for the calculations in this section\footnote{In
+  retrospect, it would likely have been nicer to have all
+  $s_\mu^{(j)}$ operators on the right throughout these calculations,
+  due to the enhanced symmetry of expressions with respect to
+  Hermitian conjugation.  In any case, we express the final result of
+  this section in both ordering conventions, and thus feel no need to
+  reproduce these calculations with a different ordering.}.  This
+choice of ordering compels us to expand
 \begin{align}
@@
   S_\mu^m S_\bmu^n
-  &= S_\mu^{m-1} S_\bmu^n S_\mu
-  + n \sp{S_\mu^{m-1} S_\bmu^{n-1} \p{1 - n + 2\mu S_\z}} \\
-  &= S_\bmu^n S_\mu^m
+  = S_\mu^{m-1} S_\bmu^n S_\mu
+  + n \sp{S_\mu^{m-1} S_\bmu^{n-1} \p{1 - n + 2\mu S_\z}}
+  = S_\bmu^n S_\mu^m
   + n \sum_{k=0}^{m-1} S_\mu^{m-k-1} S_\bmu^{n-1}
@@
 for computing the effects of single-spin decoherence on ordered
-products of collective spin operators.  These identities all involve
+products of collective spin operators, i.e.~on operators of the form
+$S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$.  These identities all involve
 sandwiching a collective spin operator between operators that act on
-individual spins only, and summing such terms over all individual spin
-indices.  Our general strategy will be to use commutation relations to
-push single-spin operators together, and then evaluate the sum to
-arrive at an expression in terms of collective spin operators only.
+individual spins only, and summing over all individual spin indices.
+Our general strategy will be to use commutation relations to push
+single-spin operators together, and then evaluate the sum to arrive at
+an expression in terms of only collective spin operators.
 
 We first compute sums of single-spin operators sandwiching
-$\p{\mu S_\z}^m$, which will in turn be used to compute similar sums
-sandwiching $S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$.  Up to Hermitian
-conjugation, the unique cases are, for $\mu,\nu\in\set{+1,-1}$,
+$\p{\mu S_\z}^m$, when necessary making use of the identity in
+\eqref{eq:push_z_mu_Ss}.  Up to Hermitian conjugation, the unique
+cases are, for $\mu,\nu\in\set{+1,-1}$,
 \begin{align}
@@
 \end{align}
-We are now equipped to derive similar identities for general ordered
-products of collective spin operators.  Making heavy use of identities
+We are now equipped to derive similar identities for more general
+collective spin operators.  Making heavy use of identities
 \eqref{eq:push_z_mu_sS} and \eqref{eq:push_mu_Ss} to push single-spin
@@
   &= S_\mu^\ell \p{S - \ell - n - \mu S_\z}
-  \p{1+\mu S_\z}^m S_\bmu^n \notag \\
-  &\qquad + \ell n \p{2S - \ell - n + 2}
+  \p{1+\mu S_\z}^m S_\bmu^n
+  + \ell n \p{2S - \ell - n + 2}
   S_\mu^{\ell-1} \p{\mu S_\z}^m S_\bmu^{n-1} \notag \\
@@
 In this section we work out the effects of permutationally-symmetric
-decoherence of individual spins on ordered products of collective spin
-operators.  For shorthand, we define
+decoherence of individual spins on collective spin operators of the
+form $S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$.  For shorthand, we define
 \begin{align}
@@
 $\gamma^{(j)}$ is an operator that acts with $\gamma$ on spin $j$ and
-the identity on all other spins.
+the identity $\1$ on all other spins.
 
@@
 \end{align}
-In order to determine the effect of this decoherence on ordered
-products of collective spin operators, we expand the anti-commutator
+In order to determine the effect of this decoherence on general
+collective spin operators, we expand the anti-commutator
 \begin{align}
@@
 $S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$ ``grow'' in $m$ through the last
-term in \eqref{eq:decay_same} with $k=m$, although the sum $\ell+m+n$
-does not grow.
+term in \eqref{eq:decay_same}, although the sum $\ell+m+n$ does not
+grow.
 
@@
 \end{align}
-In order to compute the effect of this decoherence on ordered products
-of collective spin operators, we expand the anti-commutator
+In order to compute the effect of this decoherence on general
+collective spin operators, we expand the anti-commutator
 \begin{align}
@@
 \end{align}
-Recognizing a resemblance between \eqref{eq:S_mu_acomm} and
-\eqref{eq:san_z_mu}, we can collect terms to simplify
+Recognizing a resemblance between terms in \eqref{eq:S_mu_acomm} and
+\eqref{eq:san_z_mu}, we collect terms to simplify
 \begin{align}
@@
   &\equiv -\mu n S_\mu^\ell \sp{S-\ell-\f34\p{n-1}-\f12\mu S_\z}
-  \p{\mu S_\z}^m S_\bmu^{n-1} \notag \\
-  &\qquad - \mu\ell n\p{n-1} S_\mu^{\ell-1}
+  \p{\mu S_\z}^m S_\bmu^{n-1}
+  - \mu\ell n\p{n-1} S_\mu^{\ell-1}
   \p{S+\mu S_\z} \p{-1+\mu S_\z}^m S_\bmu^{n-2}, \\
@@
   &\equiv \sum_j s_\mu^{(j)} S_\mu^\ell
-  \p{\mu S_\z}^m S_\bmu^n s_\mu^{(j)} \\
-  &= n S_\mu^{\ell+1} \p{\mu S_\z}^m S_\bmu^{n-1}
+  \p{\mu S_\z}^m S_\bmu^n s_\mu^{(j)}
+  = n S_\mu^{\ell+1} \p{\mu S_\z}^m S_\bmu^{n-1}
   - n\p{n-1} S_\mu^\ell \p{S+\mu S_\z} \p{-1+\mu S_\z}^m S_\bmu^{n-2},
@@
   = S_\bmu S_\mu \p{1+\mu S_\z}^m
-  = \p{S_\mu S_\bmu - 2\mu S_\z} \p{1+\mu S_\z}^m \\
+  = \p{S_\mu S_\bmu - 2\mu S_\z} \p{1+\mu S_\z}^m
   = S_\mu \p{2+\mu S_\z}^m S_\bmu - 2\mu S_\z \p{1+\mu S_\z}^m,
@@
 
-In this section we work out the effects of decoherence via collective
-jump operators on ordered products of collective spin operators.  For
-shorthand, we define
+In this section we work out the effects of collective decoherence on
+general collective spin operators.  For shorthand, we define
 \begin{align}
@@
 \end{align}
-where $\Gamma$ is a collective spin operator with unit weight, i.e.~a
-linear combination of $S_\z$, $S_\mu$, and $S_\bmu$.
+where $\Gamma$ is a collective spin jump operator.
 
@@
 we find that the effects of collective decay-type decoherence on
-ordered products of collective spin operators are given by
+general collective spin operators are given by
 \begin{align}
   \D\p{S_\bmu; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
-  &= -S_\mu^{\ell+1} \sp{\p{1+\mu S_\z}^m - \p{\mu S_\z}^m}
-  S_\bmu^{n+1} \notag \\
-  &\qquad + \f12 S_\mu^\ell \sp{\ell\p{\ell-1} + n\p{n-1}
+  = -S_\mu^{\ell+1} \sp{\p{1+\mu S_\z}^m - \p{\mu S_\z}^m} S_\bmu^{n+1}
+  + \f12 S_\mu^\ell \sp{\ell\p{\ell-1} + n\p{n-1}
     + 2\p{\ell+n}\mu S_\z} \p{\mu S_\z}^m S_\bmu^n,
@@
 \end{align}
-Similarly, we can work out that the effect of collective spin
-dephasing is given by
+Similarly, the effect of collective dephasing is given by
 \begin{align}
@@
 
-More generally, we may need to consider collective-spin jump operators
-of the form
+More generally, we consider jump operators of the form
 \begin{align}
@@
 \end{align}
-In order to compute the effect of this decoherence on ordered products
-of collective spin operators, we expand the anti-commutators
+In order to compute the effect of this decoherence on general
+collective spin operators, we expand the anti-commutators
 \begin{align}
@@
   &\qquad + n\p{n-1} S_\mu^\ell \sp{\p{n-1}\p{n-2}
-    + 2\p{2n-3}\mu S_\z + 4\p{\mu S_\z}^2} \notag \\
-  &\qquad\qquad\qquad\qquad \times \p{\mu S_\z}^m S_\bmu^{n-2},
+    + 2\p{2n-3}\mu S_\z + 4\p{\mu S_\z}^2} \p{\mu S_\z}^m S_\bmu^{n-2},
   \allowdisplaybreaks \\
@@
 \section{Changing bases for collective spin operators}
+\label{sec:new_basis}
 
@@
 
+\bibliography{\jobname}
+
 \end{document}

### 2018-11-28 multi_body_interactions/papers/theory/paper.tex
fixed minus sign typo, and made slight modifications to wording

@@
 single atom on lattice site $i\in\mathbb{Z}^3$ in motional state
-$n\in\mathbb{N}_0^3$ with nuclear spin $\mu\in\set{-I,I+1,\cdots,I}$
+$n\in\mathbb{N}_0^3$ with nuclear spin $\mu\in\set{-I,-I+1,\cdots,I}$
 (i.e.~projected onto a quantization axis) and orbital state
@@
   james2007effective, reiter2012effective, sanz2016adiabatic}.  Unlike
-the perturbative, but exact Schieffer-Wolff transformation, adiabatic
-elimination methods use approximations which rely on the fast dynamics
-of excited states.  While generally reasonable, these approximations
-must be made carefully to avoid potential problems with
-self-consistency (see section 3 of ref.~\cite{brion2007adiabatic}),
-and yield no obvious or straightforward means to compute effective
-corrections beyond second order in the couplings between low- and
-high-energy sectors of a Hilbert space \cite{james2007effective,
-  reiter2012effective, sanz2016adiabatic}.  While at least one attempt
-at systematically computing higher-order corrections in the framework
-of adiabatic elimination has recently been made
-\cite{sanz2016adiabatic}, the resulting expressions do not lend
-themselves as nicely to analytical treatment, and were in any case
-found by the authors to be equivalent to a Schrieffer-Wolff
-transformation.
+the perturbatively expressed, but otherwise exact Schieffer-Wolff
+transformation, adiabatic elimination methods use approximations which
+rely on the fast dynamics of excited states.  While generally
+reasonable, these approximations must be made carefully to avoid
+potential problems with self-consistency (see section 3 of
+ref.~\cite{brion2007adiabatic}), and yield no obvious or
+straightforward means to compute effective corrections beyond second
+order in the couplings between low- and high-energy sectors of a
+Hilbert space \cite{james2007effective, reiter2012effective,
+  sanz2016adiabatic}.  While at least one attempt at systematically
+computing higher-order corrections in the framework of adiabatic
+elimination has recently been made \cite{sanz2016adiabatic}, the
+resulting expressions do not lend themselves as nicely to analytical
+treatment, and were in any case found by the authors to be equivalent
+to a Schrieffer-Wolff transformation.
 
@@
 matrix in field theory.  Effective $M$-body interaction terms at order
-$p$ in $H_{\t{int}}$ can be represented by directed, weakly connected
-graphs (i.e.~which are connected when ignoring edge direction)
-containing $p$ internal vertices.  Each internal vertex and its
-associated edges correspond respectively to a coupling constant and
-the associated field operators in $H_{\t{int}}$.  An example 2-vertex
-diagram representing an effective 3-body interaction term is provided
-in figure \ref{fig:diagram}.  All diagrams are read from left to right
-to construct a sequence of operators from right to left; the internal
+$p$ in $H_{\t{int}}$ can be represented by graphs containing $p$
+internal vertices.  Each internal vertex and its associated edges
+correspond respectively to a coupling constant and the associated
+field operators in $H_{\t{int}}$.  An example 2-vertex diagram
+representing an effective 3-body interaction term is provided in
+figure \ref{fig:diagram}.  All diagrams are read from left to right to
+construct a sequence of operators from right to left; the internal
 vertices of a diagram are thus strictly ordered, with the $n$-th

### 2018-11-29 soc_squeezing/paper/collective_spin.tex
added weak decoherence figure caption and made minor modifications to text

@@
   \includegraphics{coherent.pdf}
-  \caption{Spin squeezing of $N=10^4$ spins under coherent evolution,
-    as computed using exact methods (solid lines) and the expansion in
+  \caption{Spin squeezing of $N=10^4$ spins initially in $\ket\X$
+    under coherent evolution, as computed using exact methods (solid
+    lines) and the expansion in Eq.~\eqref{eq:order_truncation} with
+    $M=40$ (dashed lines).  Solid circles mark the times for TAT and
+    TNT after which the latter method gives an unphysical result with
+    $\xi^2<0$.  Both methods agree exactly in the OAT case for all
+    times shown.}
+  \label{fig:coherent}
+\end{figure}
+
+\begin{figure}
+  \centering
+  % \includegraphics{weak_decoherence.pdf}
+  \caption{Spin squeezing of $N=10^4$ spins initially in $\ket\X$ with
+    decoherence via uncorrelated decay, excitation, and dephasing of
+    individual spins at rates $\gamma_-=\gamma_+=\gamma_\z=\chi$.
+    Squeezing is computed using an exact simulation (solid line),
+    quantum trajectory methods (dots), and the expansion in
     Eq.~\eqref{eq:order_truncation} with $M=40$ (dashed lines).  Solid
     circles mark the times for TAT and TNT after which the latter
-    method gives an unphysical result with $\xi^2<0$.  Both methods
-    agree exactly in the OAT case for all times shown.}
-  \label{fig:coherent}
+    method gives an unphysical result with $\xi^2<0$.}
+  \label{fig:weak_decoherence}
 \end{figure}
@@
 operators $\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$ contain
-genuinely $\abs{\v m}$-body operators (for
-$\abs{\v m}\equiv\sum_\mu m_\mu$), but also carry lots of ``baggage''
-in the form of $k$-body operators with $k<\abs{\v m}$.  To cut down on
-this overhead, we try to change our basis for collective spin
-operators from $\S_{\v m}$ to the purely-$\abs{\v m}$-body operators
+$\abs{\v m}$-body operators (for $\abs{\v m}\equiv\sum_\mu m_\mu$),
+but also carry lots of ``baggage'' in the form of $k$-body operators
+with $k<\abs{\v m}$.  To cut down on this overhead, we try to change
+our basis for collective spin operators from $\S_{\v m}$ to the
+purely-$\abs{\v m}$-body operators
 \begin{align}
@@
 $\abs{\set{j^\mu_a}} = m_\mu$, such that $j^\mu_a$ indexes the $a$-th
-spin addressed by a $s_\mu$ operator in $P_{\v j}$.
+spin addressed by an $s_\mu$ operator in $P_{\v j}$.
 
@@
 \end{align}
-where $\simeq$ denotes equality up to a re-indexing of spins; we have
-introduced explicit dependence on
+where $\simeq$ denotes equality up to a re-indexing of spins and
+additional tensor factors of the single-spin identity operator
+(i.e.~$\1$); we have introduced explicit dependence on
 $r_{\mu\nu}=\abs{\set{K^{\mu\nu}_a}}$ for brevity; and
@@
 $\rho^{\mu\nu}_a$ indexes one of the terms in that factor.  Letting
-$\tilde\rho^{\mu\nu}_\kappa$ denote the number of elements in
+$\tilde\rho^{\mu\nu}_\kappa$ denote the number of elements in the
+index list
 $\p{\rho^{\mu\nu}_1,\rho^{\mu\nu}_2,\cdots,\rho^{\mu\nu}_{r_{\mu\nu}}}$
@@
 \begin{align}
-  \max\set{0,\sum_\mu\p{m_\mu+n_\mu}-N}
-  \le s \le \min\set{N,\sum_\mu\p{m_\mu+n_\mu}}.
+  \max\set{0,\abs{\v m}+\abs{\v n}-N}
+  \le s \le \min\set{N,\abs{\v m}+\abs{\v n}}.
 \end{align}
@@
 all $\mu,\nu$ with $s_\mu s_\nu=0$.  Finally, we can avoid the
-nullifying effect of zeros from $\eta_{\mu\nu\kappa}$
-$\v\eta_{\v\rho}\propto\p{\eta_{\mu\nu\kappa}}^{\rho^{\mu\nu}_\kappa}$
-by fixing $\rho^{\mu\nu}_\kappa=0$ for all $\mu,\nu,\kappa$ with
+nullifying effect of zeros from
+$\v\eta_{\v\rho}=\prod_{\mu,\nu,\kappa}
+\p{\eta_{\mu\nu\kappa}}^{\rho^{\mu\nu}_\kappa}$ by fixing
+$\rho^{\mu\nu}_\kappa=0$ for all $\mu,\nu,\kappa$ with
 $\eta_{\mu\nu\kappa}=0$.  These restrictions are sufficient to ensure

### 2018-12-03 soc_squeezing/paper/collective_spin.tex
removed obsolete text
modified first intro paragraph, added some references

@@
 
-Collective spin systems attract considerable attention for providing a
-window into many-body quantum phenomena that are both experimentally
-accessible\cite{takano2009spin, chen2010heralded,
-  riedel2010atomchipbased, gross2010nonlinear, martin2013quantum,
-  bohnet2016quantum, hosten2016quantum, norcia2018cavitymediated,
-  jones2009magnetic} and metrologically
-useful\cite{riedel2010atomchipbased, gross2010nonlinear,
-  hosten2016quantum, jones2009magnetic, zhong2010simplified,
-  huang2015quantum, agarwal1997atomic, lau2014proposal,
-  wineland1992spin, kitagawa1993squeezed, ma2011quantum}.  Theoretical
-interest in such systems dates back to the mid-twentieth century with
-the introduction of the Lipkin-Meshkov-Glick (LMG) model as a toy
-model for testing many-body approximation methods in contemporary
-nuclear physics\cite{lipkin1965validity, meshkov1965validity,
+Collective spin systems are a versatile resource in quantum science
+for a range of applications including quantum-enhanced metrology,
+quantum simulation, and quantum information.  Theoretical interest in
+such systems dates back to the mid-twentieth century with the
+introduction of the Lipkin-Meshkov-Glick (LMG) model as a toy model
+for testing many-body approximation methods in contemporary nuclear
+physics\cite{lipkin1965validity, meshkov1965validity,
   glick1965validity}.  Since then, collective system systems have
-additionally featured in theoretical studies of quantum
-criticality\cite{latorre2005entanglement, alcalde2007functional,
+additionally featured in both theoretical and experimental studies of
+quantum criticality\cite{latorre2005entanglement,
+  alcalde2007functional, morrison2008dissipationdriven,
   morrison2008collective, sarandy2009classical, wang2012quantum,
-  majd2014lmg}, non-equilibrium
-phenomena\cite{morrison2008dissipationdriven,
+  kessler2012dissipative, majd2014lmg}, non-equilibrium
+phenomena\cite{walls1978nonequilibrium, morrison2008dynamical,
   bhattacherjee2014nonequilibrium, klinder2015dynamical,
-  zhiqiang2017nonequilibrium, smale2018observation}, and precision
+  maghrebi2016nonequilibrium, zhiqiang2017nonequilibrium,
+  lang2018concurrence, smale2018observation}, and precision
 metrology\cite{wineland1992spin, kitagawa1993squeezed, ma2011quantum,
@@
 ensembles\cite{takano2009spin, chen2010heralded}, Bose-Einstein
-condensates\cite{riedel2010atomchipbased, gross2010nonlinear},
-ultracold Fermi gasses\cite{martin2013quantum}, trapped
+condensates\cite{riedel2010atomchipbased, gross2010nonlinear,
+  klinder2015dynamical}, ultracold Fermi
+gasses\cite{martin2013quantum, smale2018observation}, trapped
 ions\cite{bohnet2016quantum}, optical cavities\cite{hosten2016quantum,
+  zhiqiang2017nonequilibrium, lewis-swan2018robust,
   norcia2018cavitymediated}, and molecular
@@
 such as NOON\cite{jones2009magnetic, chen2010heralded,
-  zhong2010simplified}, spin-cat\cite{huang2015quantum,
-  agarwal1997atomic, lau2014proposal}, and spin-squeezed
+  zhong2010simplified}, spin-cat\cite{agarwal1997atomic,
+  lau2014proposal, huang2015quantum}, and spin-squeezed
 states\cite{takano2009spin, riedel2010atomchipbased,
-  gross2010nonlinear, hosten2016quantum, norcia2018cavitymediated,
-  wineland1992spin, kitagawa1993squeezed, ma2011quantum} that allow
-for measurement errors $\Delta\phi\sim1/N^\epsilon$ with
-$\epsilon>1/2$.  Generating such states requires nonlinear dynamics
-that are typically realized using long-range collisional, photon-, or
-phonon-mediated interactions.
-% typically resulting in a one-axis twising (OAT) model described by
-% the Hamiltonian\cite{kitagawa1993squeezed}
-% \begin{align}
-%   H_{\t{OAT}} = \chi S_\z^2,
-%   \label{eq:OAT}
-% \end{align}
-% where $S_\z\equiv\sum_{j=1}^N\sigma_\z^{(j)}/2$ is a collective
-% spin-$z$ operator defined in terms of Pauli-$z$ operators
-% $\sigma_\z^{(j)}$ addressing spin $j$.  The OAT model prepares a
-% spin-squeezed state that allows measurement of a phase angle $\phi$
-% with error $\Delta\phi\sim N^{-2/3}$ at times
-% $t\sim\chi^{-1}N^{-2/3}$\cite{ma2011quantum}.
+  gross2010nonlinear, hosten2016quantum, lewis-swan2018robust,
+  norcia2018cavitymediated, wineland1992spin, kitagawa1993squeezed,
+  ma2011quantum} that allow for measurement errors
+$\Delta\phi\sim1/N^\epsilon$ with $\epsilon>1/2$.  Generating such
+states requires nonlinear dynamics that are typically realized using
+long-range collisional, photon-, or phonon-mediated interactions.
 Although a truly collective spin model requires uniform, all-to-all
@@
 parameters\cite{hu2015entangled}.
-% Various works have proposed improvements over OAT through the
-% addition of auxiliary control fields\cite{law2001coherent,
-% liu2011spin, huang2015twoaxis, muessel2015twistandturn} or a
-% modified vacuum\cite{hu2017vacuum}, thereby realizing different
-% collective spin models.
-
-% While the OAT model admits exact
-% solutions\cite{foss-feig2013nonequilibrium}, the same is not true of
-% more general, non-Ising Hamiltonians.  Nonetheless,
+
 In the absence of decoherence, permutation symmetry and total spin
@@
 $N\lesssim100$ particles.  If decoherence is sufficiently weak,
-dynamics are numerically solvable for $N\lesssim10^4$ particles via
+dynamics can be numerically solvable for $N\lesssim10^5$ particles via
 ``quantum trajectory'' Monte Carlo methods\cite{plenio1998quantumjump,
@@
 discuss briefly at the end of Appendix \ref{sec:new_basis}.  As we
-will see in the following sections, our choice of the basis
-$\set{\S_{\v m}}$ allows us to accurately compute spin correlators to
-times that are relevant e.g.~for spin squeezing protocols.
+will see in the following sections, our choice of basis elements
+$\S_{\v m}=S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$ allows us to accurately
+compute spin correlators to times that are relevant e.g.~for spin
+squeezing protocols.
 

### 2018-12-04 soc_squeezing/paper/collective_spin.tex
made first sentence of intro more broad added more references to introductory paragraphs

@@
   lang2018concurrence, smale2018observation}, and precision
-metrology\cite{wineland1992spin, kitagawa1993squeezed, ma2011quantum,
-  zhong2010simplified}.  With the development of advanced atomic
-trapping, cooling, and control techniques, collective spin systems
-have been experimentally realized in atomic
-ensembles\cite{takano2009spin, chen2010heralded}, Bose-Einstein
+metrology\cite{wineland1992spin, itano1993quantum,
+  kitagawa1993squeezed, ma2011quantum, chen2010heralded,
+  takano2009spin, appel2009mesoscopic, schleier-smith2010states,
+  chen2011conditional, riedel2010atomchipbased, gross2010nonlinear,
+  zhong2010simplified, agarwal1997atomic, lau2014proposal,
+  huang2015quantum, wineland1992spin, kitagawa1993squeezed,
+  schleier-smith2010squeezing, leroux2010implementation,
+  bohnet2014reduced, cox2016deterministic, hosten2016measurement,
+  hosten2016quantum, lewis-swan2018robust, norcia2018cavitymediated,
+  jones2009magnetic}.  With the development of advanced AMO (atomic,
+molecular, and optical) trapping, cooling, and control techniques,
+collective spin systems have been experimentally realized in atomic
+ensembles\cite{takano2009spin, appel2009mesoscopic, chen2010heralded,
+  schleier-smith2010states, chen2011conditional}, Bose-Einstein
 condensates\cite{riedel2010atomchipbased, gross2010nonlinear,
@@
 gasses\cite{martin2013quantum, smale2018observation}, trapped
-ions\cite{bohnet2016quantum}, optical cavities\cite{hosten2016quantum,
+ions\cite{bohnet2016quantum}, optical
+cavities\cite{leroux2010implementation, bohnet2014reduced,
+  cox2016deterministic, hosten2016measurement, hosten2016quantum,
   zhiqiang2017nonequilibrium, lewis-swan2018robust,
@@
   lau2014proposal, huang2015quantum}, and spin-squeezed
-states\cite{takano2009spin, riedel2010atomchipbased,
-  gross2010nonlinear, hosten2016quantum, lewis-swan2018robust,
-  norcia2018cavitymediated, wineland1992spin, kitagawa1993squeezed,
+states\cite{wineland1992spin, kitagawa1993squeezed, takano2009spin,
+  appel2009mesoscopic, schleier-smith2010states,
+  schleier-smith2010squeezing, leroux2010implementation,
+  riedel2010atomchipbased, gross2010nonlinear, chen2011conditional,
+  bohnet2014reduced, cox2016deterministic, hosten2016measurement,
+  hosten2016quantum, lewis-swan2018robust, norcia2018cavitymediated,
   ma2011quantum} that allow for measurement errors
 $\Delta\phi\sim1/N^\epsilon$ with $\epsilon>1/2$.  Generating such
-states requires nonlinear dynamics that are typically realized using
+states typically requires nonlinear dynamics that are realized using
 long-range collisional, photon-, or phonon-mediated interactions.

### 2018-12-11 soc_squeezing/paper/collective_spin.tex
updated text with Ana Maria's edits

@@
 such systems dates back to the mid-twentieth century with the
-introduction of the Lipkin-Meshkov-Glick (LMG) model as a toy model
+introduction of the Dicke model\cite{dicke1954coherence} that
+describes atoms cooperatively interacting with a single mode of the
+radiation field and the Lipkin-Meshkov-Glick (LMG) model, a toy model
 for testing many-body approximation methods in contemporary nuclear
 physics\cite{lipkin1965validity, meshkov1965validity,
-  glick1965validity}.  Since then, collective system systems have
-additionally featured in both theoretical and experimental studies of
-quantum criticality\cite{latorre2005entanglement,
-  alcalde2007functional, morrison2008dissipationdriven,
-  morrison2008collective, sarandy2009classical, wang2012quantum,
-  kessler2012dissipative, majd2014lmg}, non-equilibrium
-phenomena\cite{walls1978nonequilibrium, morrison2008dynamical,
-  bhattacherjee2014nonequilibrium, klinder2015dynamical,
-  maghrebi2016nonequilibrium, zhiqiang2017nonequilibrium,
-  lang2018concurrence, smale2018observation}, and precision
-metrology\cite{wineland1992spin, itano1993quantum,
-  kitagawa1993squeezed, ma2011quantum, chen2010heralded,
-  takano2009spin, appel2009mesoscopic, schleier-smith2010states,
-  chen2011conditional, riedel2010atomchipbased, gross2010nonlinear,
-  zhong2010simplified, agarwal1997atomic, lau2014proposal,
-  huang2015quantum, wineland1992spin, kitagawa1993squeezed,
-  schleier-smith2010squeezing, leroux2010implementation,
-  bohnet2014reduced, cox2016deterministic, hosten2016measurement,
-  hosten2016quantum, lewis-swan2018robust, norcia2018cavitymediated,
-  jones2009magnetic}.  With the development of advanced AMO (atomic,
-molecular, and optical) trapping, cooling, and control techniques,
-collective spin systems have been experimentally realized in atomic
+  glick1965validity}.  On the experimental side, the development of
+advanced trapping, cooling, and control techniques in atomic,
+molecular, and optical (AMO) systems have enabled the realization of
+collective spin models in a broad range of platforms, including atomic
 ensembles\cite{takano2009spin, appel2009mesoscopic, chen2010heralded,
@@
   norcia2018cavitymediated}, and molecular
-sensors\cite{jones2009magnetic}.
+sensors\cite{jones2009magnetic}. These implementations have triggered
+intensive investigations of rich phenomena, including studies of
+quantum criticality\cite{latorre2005entanglement,
+  alcalde2007functional, morrison2008dissipationdriven,
+  morrison2008collective, sarandy2009classical, wang2012quantum,
+  kessler2012dissipative, majd2014lmg}, non-equilibrium
+phenomena\cite{walls1978nonequilibrium, morrison2008dynamical,
+  bhattacherjee2014nonequilibrium, klinder2015dynamical,
+  maghrebi2016nonequilibrium, zhiqiang2017nonequilibrium,
+  lang2018concurrence, smale2018observation}, and precision
+metrology\cite{wineland1992spin, itano1993quantum,
+  kitagawa1993squeezed, ma2011quantum, takano2009spin,
+  appel2009mesoscopic, chen2010heralded, schleier-smith2010states,
+  chen2011conditional, riedel2010atomchipbased, gross2010nonlinear,
+  zhong2010simplified, agarwal1997atomic, lau2014proposal,
+  huang2015quantum, wineland1992spin, kitagawa1993squeezed,
+  schleier-smith2010squeezing, leroux2010implementation,
+  bohnet2014reduced, cox2016deterministic, hosten2016measurement,
+  hosten2016quantum, lewis-swan2018robust, norcia2018cavitymediated,
+  jones2009magnetic}.
 
@@
 long time to converge, as simulations become dominated by incoherent
-jumps that generate large numbers of distinct trajectories over which
-to average.
+jumps that generate large numbers of distinct trajectories that need
+to be averaged.
 
@@
 critical value known to maximize the entanglement generation rate of
-of TNT in the large-$N$ limit\cite{micheli2003manyparticle,
-  sorelli2015fast}.  Note that the OAT model is a special case of the
-zero-field Ising model, whose quantum dynamics in the presence of
-decoherence via uncorrelated decay, excitation, and dephasing of
-individual spins have been previously solved
-analytically\cite{foss-feig2013nonequilibrium} (see Appendix
-\ref{sec:OAT}, as well as the Supplementary Material of
-Ref.~[\citenum{bohnet2016quantum}] for more general multi-spin
-correlators).  We will benchmark against these analytical OAT results
-wherever applicable.
+TNT in the large-$N$ limit\cite{micheli2003manyparticle,
+  sorelli2015fast}.
+
+Note that the OAT model is a special case of the zero-field Ising
+model, whose quantum dynamics admits an exact analytic solution even
+in the presence of decoherence\cite{foss-feig2013nonequilibrium}.  We
+will benchmark our calculations using these analytical OAT results
+wherever applicable (see Appendix \ref{sec:OAT}, as well as the
+Supplementary Material of Ref.~[\citenum{bohnet2016quantum}]).
 
@@
 with $S_\x\ket{\X}=S\ket{\X}$.  Our measure of spin squeezing is the
-directionally-unbiased squeezing parameter determined by the maximal
-gain in resolution $\Delta\phi$ of a phase angle $\phi$ over that
-achieved by any spin-polarized product state
+directionally-unbiased Ramsey squeezing parameter determined by the
+maximal gain in resolution $\Delta\phi$ of a phase angle $\phi$ over
+that achieved by any spin-polarized product state
 (e.g.~$\ket{\X}$)\cite{wineland1992spin, ma2011quantum},
@@
 the form $\bk{S_m}$ and $\bk{S_m S_n}$.  In the case of coherent
-dynamics under TAT or TNT, these correlators are obtainable via exact
+dynamics discussed here, these correlators are obtainable via exact
 simulations of quantum dynamics in the $\p{N+1}$-dimensional Dicke
@@
 $\bk{S,m|\v S^2|S,m}=S\p{S+1}$ and $\bk{S,m|S_\z|S,m}=m$ for
-$\abs{m}\le S$.  In the presence of permutationally-symmetric
-decoherence, these correlators are obtainable with the collective-spin
-quantum trajectory Monte Carlo method recently developed in
-ref.~\cite{zhang2018montecarlo}.  These exact and quantum trajectory
-simulations will serve as our methods for benchmarking the expansion
-in Eq.~\eqref{eq:order_truncation}.
+$m\in\set{-S,-S+1,\cdots,S}$.  In the presence of
+permutationally-symmetric decoherence, these correlators are
+obtainable with the collective-spin quantum trajectory Monte Carlo
+method recently developed in ref.~\cite{zhang2018montecarlo}.  In this
+work, these exact and quantum trajectory simulations will be used to
+benchmark the expansion in Eq.~\eqref{eq:order_truncation}.
 
@@
     lines) and the expansion in Eq.~\eqref{eq:order_truncation} with
-    $M=40$ (dashed lines).  Solid circles mark the times for TAT and
-    TNT after which the latter method gives an unphysical result with
-    $\xi^2<0$.  Both methods agree exactly in the OAT case for all
-    times shown.}
+    $M=40$ (dashed lines).  Solid circles mark the times at which the
+    latter method gives an unphysical result with $\xi^2<0$.  Both
+    methods agree exactly in the OAT case for all times shown.}
   \label{fig:coherent}
@@
     Eq.~\eqref{eq:order_truncation} with $M=40$ (dashed lines).  Solid
-    circles mark the times for TAT and TNT after which the latter
-    method gives an unphysical result with $\xi^2<0$.}
+    circles mark the times at which the latter method gives an
+    unphysical result with $\xi^2<0$.}
   \label{fig:weak_decoherence}
@@
 \end{align}
-where ${ n \brack k }$ is an unsigned Stirling number of the first
+where ${ p \brack q }$ is an unsigned Stirling number of the first
 kind, and finally
@@
   \p{-\sp{m-p} + \mu S_\z}^\ell
-  = \sum_{\ell=0}^p \epsilon_{mn}^{p\ell}
-  \p{-1}^\ell \p{\sp{m-p} + \bmu S_\z}^\ell
   = \sum_{q=0}^p \zeta_{mn}^{pq} \p{\bmu S_\z}^q,
@@
   \zeta_{mn}^{pq}
-  \equiv \sum_{\ell=q}^p \epsilon_{mn}^{p\ell} \p{-1}^\ell
-  { \ell \choose q } \p{m-p}^{\ell-q}
+  \equiv \sum_{\ell=q}^p \epsilon_{mn}^{p\ell}
+  { \ell \choose q } \p{-1}^\ell \p{m-p}^{\ell-q}
   = \p{-1}^p 2^q \sum_{s=q}^p
@@
 \end{align}
+Here ${ p \brack s }$ is an unsigned Stirling number of the first
+kind.
 

### 2019-01-17 multi_body_interactions/papers/theory/paper.tex
make some subscripts normal (rather than math) font

@@
 \newcommand{\bk}{\braket} % shorthand for braket
+
 \newcommand{\g}{\text{g}}
 \newcommand{\e}{\text{e}}
+\newcommand{\x}{\text{x}}
+\newcommand{\y}{\text{y}}
+\newcommand{\z}{\text{z}}
+
 \renewcommand{\c}{\hat{c}}
 \newcommand{\n}{\hat{n}}
+
 \newcommand{\A}{\mathcal{A}}
@@
 motional state $n$.  In a harmonic trap approximation we would have
-$E_n=\p{3/2+n_x+n_y+n_z}\omega$ for an on-site angular trap frequency
-$\omega$ (with the reduced Planck constant $\hbar=1$ throughout this
-paper), but in general the aharmonicity of the lattice potential will
-cause a non-negligible shift in motional state energies.
+$E_n=\p{3/2+n_\x+n_\y+n_\z}\omega$ for an on-site angular trap
+frequency $\omega$ (with the reduced Planck constant $\hbar=1$
+throughout this paper), but in general the aharmonicity of the lattice
+potential will cause a non-negligible shift in motional state
+energies.
 

### 2019-01-29 multi_body_interactions/papers/theory/permissions.tex
added document explaning permissions to reproduce published materials

\documentclass[nofootinbib,notitlepage,11pt]{revtex4-1}

%%% header / footer
\usepackage{fancyhdr} % easier header and footer management
\pagestyle{fancy} % page formatting style
\fancyhf{}
\renewcommand{\headrulewidth}{0pt} % horizontal line in header

\usepackage{hyperref,listings}
\setlength{\parindent}{0pt}
\parskip 6pt

\begin{document}

\title{Concerning permissions for reproducing material}

\maketitle
\thispagestyle{fancy}

The manuscript here submitted to {\it New Journal of Physics} contains
two figures which are reproduced from another manuscript currently in
press (accepted, but not yet published) with {\it Nature}, titled
``Emergence of multi-body interactions in a fermionic lattice clock''
(DOI: \verb|10.1038/s41586-018-0661-6|).  Both authors of this
manuscript are also authors of the one in {\it Nature}, and grant {\it
  New Journal of Physics} permission to use the reproduced material in
question.  Furthermore, it is editorial policy of Springer Nature, the
publishing company that owns {\it Nature}, that ownership of copyright
remains with the authors of its published material.  Verbatim from the
\href{https://www.nature.com/reprints/permission-requests.html}{Springer
  Nature permission request webpage}:
\begin{lstlisting}[breaklines]
  The author of articles published by SpringerNature do not usually
  need to seek permission for re-use of their material as long as the
  journal is credited with initial publication.

  Ownership of copyright in in original research articles remains with
  the Author
\end{lstlisting}
The publisher-to-be of the reproduced material thus also implicitly
grants permission to use it, conditional on proper citation practices.
As the reproduced material is currently in press and not yet
published, the manuscript submitted to {\it New Journal of Physics}
cites the arXiv version of the original material
(\href{https://arxiv.org/abs/1803.11282}{arXiv:1803.11282}).  Once the
original material is published, this citation will be updated
accordingly.

If we have in any way misunderstood editorial policy of either
Springer Nature or the Institute of Physics, please let us know and we
will take whatever actions necessary to sort out issues relating to
reproduced content and copyright.  In the meantime, we hope the
editors of {\it New Journal of Physics} can proceed to consider our
submitted manuscript without delay.

~\\

Sincerely,

Michael A. Perlin

Ana Maria Rey

\end{document}

### 2019-01-29 multi_body_interactions/papers/theory/paper.tex
modified wording in text to minimize the number of revisions that are not made in response to referee comments

@@
   james2007effective, reiter2012effective, sanz2016adiabatic}.  Unlike
-the perturbatively expressed, but otherwise exact Schieffer-Wolff
-transformation, adiabatic elimination methods use approximations which
-rely on the fast dynamics of excited states.  While generally
-reasonable, these approximations must be made carefully to avoid
-potential problems with self-consistency (see section 3 of
-ref.~\cite{brion2007adiabatic}), and yield no obvious or
-straightforward means to compute effective corrections beyond second
-order in the couplings between low- and high-energy sectors of a
-Hilbert space \cite{james2007effective, reiter2012effective,
-  sanz2016adiabatic}.  While at least one attempt at systematically
-computing higher-order corrections in the framework of adiabatic
-elimination has recently been made \cite{sanz2016adiabatic}, the
-resulting expressions do not lend themselves as nicely to analytical
-treatment, and were in any case found by the authors to be equivalent
-to a Schrieffer-Wolff transformation.
+the perturbative, but exact Schieffer-Wolff transformation, adiabatic
+elimination methods use approximations which rely on the fast dynamics
+of excited states.  While generally reasonable, these approximations
+must be made carefully to avoid potential problems with
+self-consistency (see section 3 of ref.~\cite{brion2007adiabatic}),
+and yield no obvious or straightforward means to compute effective
+corrections beyond second order in the couplings between low- and
+high-energy sectors of a Hilbert space \cite{james2007effective,
+  reiter2012effective, sanz2016adiabatic}.  While at least one attempt
+at systematically computing higher-order corrections in the framework
+of adiabatic elimination has recently been made
+\cite{sanz2016adiabatic}, the resulting expressions do not lend
+themselves as nicely to analytical treatment, and were in any case
+found by the authors to be equivalent to a Schrieffer-Wolff
+transformation.
 
@@
 matrix in field theory.  Effective $M$-body interaction terms at order
-$p$ in $H_{\t{int}}$ can be represented by graphs containing $p$
-internal vertices.  Each internal vertex and its associated edges
+$p$ in $H_{\t{int}}$ can be represented by directed graphs containing
+$p$ internal vertices.  Each internal vertex and its associated edges
 correspond respectively to a coupling constant and the associated

### 2019-02-02 multi_body_interactions/papers/theory/paper.tex
updated multi-body paper with edits in response to referee comments

@@
 
+\newcommand{\mA}{m_{\text{A}}} % symbol for the mass of an atom
+
 % "left vector" arrow; requires tikz package
@@
 
+\usepackage{setspace} % to modify spacing in table of contents
+
 % leave a note in the text, visible in the compiled document
@@
   the prospect of performing quantum simulations of spin models and
-  lattice field theories exhibiting SU($N$) symmetry.  Motivated in
-  particular by recent experiments preparing high density samples of
+  lattice field theories exhibiting SU($N$) symmetry.  In order to
+  analyze recent experiments preparing high density samples of
   strongly interacting ${}^{87}$Sr atoms in a three-dimensional
@@
 
+\begin{spacing}{1.3}
   \tableofcontents
+\end{spacing}
 
@@
 \label{sec:introduction}
-
-% Rapid advancements in the experimental capabilities with ultracold
-% alkaline-earth-like atoms (AEAs) \cite{fukuhara2007degenerate,
-% stellmer2009boseeinstein, deescobar2009boseeinstein,
-% tey2010doubledegenerate, desalvo2010degenerate, taie2012su,
-% martin2013quantum, cappellini2014direct, scazza2014observation,
-% zhang2014spectroscopic, cazalilla2014ultracold} bring to a
-% surprisingly near term the prospect of performing quantum
-% simulations of spin models and lattice field theories exhibiting
-% SU($N$) symmetry \cite{cazalilla2009ultracold, hermele2009mott,
-% hermele2011topological, gorshkov2010twoorbital, banerjee2013atomic,
-% cazalilla2014ultracold}.  AEAs have ground and metastable
-% (electronic) orbital states\note{[I want to briefly discuss how we
-% present this]} ${}^1S_0$ and ${}^3P_0$, often referred to as the
-% ``clock'' states due to their use in optical lattice clocks
-% \cite{derevianko2011colloquium, katori2011optical}, with
-% ultra-narrow (millihertz) optical transition linewidths and
-% associated $\sim10^2$ second excited-state lifetimes.  Due to the
-% lack of electronic orbital angular momentum in the ${}^1S_0$ and
-% ${}^3P_0$ states, AEAs exhibit decoupled orbital and nuclear spin
-% degrees of freedom.  For atoms with total nuclear spin $I$ and
-% $N=2I+1$ nuclear spin levels, this decoupling leads to
-% nuclear-spin-conserving SU($N$)-symmetric interactions governed
-% entirely by orbital-state parameters \cite{cazalilla2009ultracold,
-% taie2012su, zhang2014spectroscopic}.
 
@@
 achieved a total fractional uncertainty of $2\times10^{-18}$
-\cite{bloom2014optical, nicholson2015systematic}.  Last year (2017), a
-new generation of OLCs became operational at JILA, interrogating a
-Fermi degenerate gas of ${}^{87}$Sr atoms in a 3-D lattice at
-nanokelvin temperatures \cite{campbell2017fermidegenerate}.  All of
-these atoms' degrees of freedom, including the electronic orbital,
-nuclear spin, and motional states, can be fully controlled with high
-fidelity in a 3-D lattice \cite{daley2008quantum,
+\cite{bloom2014optical, nicholson2015systematic}.  More recently
+(2017), a new generation of OLCs became operational at JILA,
+interrogating a Fermi degenerate gas of ${}^{87}$Sr atoms in a 3-D
+lattice at nanokelvin temperatures \cite{campbell2017fermidegenerate}.
+All of these atoms' degrees of freedom, including the electronic
+orbital, nuclear spin, and motional states, can be fully controlled
+with high fidelity in a 3-D lattice \cite{daley2008quantum,
   gorshkov2009alkalineearthmetal, daley2011quantum,
@@
 
-% One of the major drivers of experimental capabilities with AEAs is
-% the development of optical lattice clocks
-% \cite{derevianko2011colloquium, katori2011optical}.  These clocks
-% traditionally operate in a one-dimensional (1-D) geometry with
-% harmonic confinement of atoms in the plane orthogonal to the lattice
-% axis \cite{derevianko2011colloquium, hinkley2013atomic,
-% bloom2014optical}.  By operating at higher atomic densities, one can
-% simultaneously interrogate a larger number of atoms and thereby
-% increase the signal-to-noise ratio of the clock.  With the extreme
-% sensitivities accessible to state-of-the-art optical lattice clocks,
-% however, 1-D clock experiments with ${}^{87}$Sr and ${}^{171}$Yb
-% detect density-dependent fractional frequency shifts at the
-% $10^{-16}$ level \cite{campbell2009probing, lemke2009spin}.  For the
-% purpose of timekeeping, these frequency shifts amount to an
-% undesirable source of systematic error, but they introduced a new,
-% exciting paradigm for the field of quantum simulation: a marriage
-% between precision clock spectroscopy and quantum many-body physics
-% \cite{taie2012su, scazza2014observation, martin2013quantum,
-% cappellini2014direct, zhang2014spectroscopic, rey2014probing}.
-% Somewhat counter-intuitively, one of the most promising strategies
-% for improving the interaction-limited 1-D clock design has been to
-% use higher dimensional lattices to generate tighter on-site
-% confinement of atoms, thereby {\it increasing} the strength of
-% inter-atomic interactions and making evolution into interacting
-% many-body states energetically unfavorable
-% \cite{swallows2011suppression, lemke2011wave,
-% campbell2017fermidegenerate}.  A wonderful side-effect of this
-% strategy for building better clocks is the development of a highly
-% controllable quantum many-body system in the strongly-interacting
-% regime, where inter-atomic interactions set the largest energy scale
-% relevant for system dynamics.
-
 In this work, we investigate the first experimental capabilities with
@@
 lattice with multiple occupation of individual lattice sites
-\cite{goban2018emergence}.  Specifically, we consider the preparation
-of isolated few-body systems in the deep-lattice limit, and carry out
-a bottom-up investigation of emergent multi-body interactions on
-multiply-occupied lattice sites.  These multi-body interactions appear
-in a low-energy effective theory of the atoms, and inherit the SU($N$)
-symmetry of their bare, pair-wise interactions, thereby enabling
-experimental studies of multi-body SU($N$) physics through the
-exquisite capabilities with OLCs.
+\cite{goban2018emergence}.  Specifically, we consider ground-state
+preparation of isolated few-body systems in the deep-lattice limit,
+and carry out a bottom-up investigation of emergent multi-body
+interactions on multiply-occupied lattice sites.  These multi-body
+interactions appear in a low-energy effective theory of the atoms, and
+inherit the SU($N$) symmetry of their bare, pair-wise interactions,
+thereby enabling experimental studies of multi-body SU($N$) physics
+through the exquisite capabilities with OLCs.  We briefly summarize
+the experimental procedures relevant to our work in Appendix
+\ref{sec:experiment_summary}.
 
@@
 In the absence of hyperfine coupling, as when addressing the spinless
-${}^1S_0$ ($\g$) and ${}^3P_0$ ($\e$) orbital states, interactions
-between any two atoms are governed by their orbital states alone, and
-are therefore characterized by four scattering lengths $a_X$ with
-$X\in\set{\g\g,\e\g^-,\e\g^+,\e\e}$, where the $+$ ($-$) superscript
-denotes symmetrization (anti-symmetrization) of a two-body orbital
-state under particle exchange.  In the low-energy limit, we can write
-the bare two-body interaction Hamiltonian in the form
-\cite{gorshkov2010twoorbital}
+${}^1S_0$ ($\g$) and ${}^3P_0$ ($\e$) orbital states of AEAs,
+interactions between any two atoms are governed by their orbital
+states alone, and are therefore characterized by four scattering
+lengths $a_X$ with $X\in\set{\g\g,\e\g^-,\e\g^+,\e\e}$, where the $+$
+($-$) superscript denotes symmetrization (anti-symmetrization) of a
+two-body orbital state under particle exchange.  In the low-energy
+limit, we can write the bare two-body interaction Hamiltonian in the
+form \cite{gorshkov2010twoorbital}
 \begin{align}
@@
   \hat\psi_{\nu,\e} \hat\psi_{\mu,\g},
+  \label{eq:H_int_start}
 \end{align}
@@
 \begin{align}
-  G_{s=\g,\e} \equiv \f{4\pi}{m_A}~ a_{ss}, &&
-  G_\pm \equiv \f{2\pi}{m_A}~ \p{a_{\e\g+} \pm a_{\e\g-}},
+  G_{s=\g,\e} \equiv \f{4\pi}{\mA}~ a_{ss}, &&
+  G_\pm \equiv \f{2\pi}{\mA}~ \p{a_{\e\g+} \pm a_{\e\g-}},
   \label{eq:couplings}
 \end{align}
-where $m_A$ is the mass of a single atom.  Defining for brevity
+where $\mA$ is the mass of a single atom.  Defining for brevity
 \begin{align}
@@
 coupling constant between the two-atom states
-$(\mu,q)+(\nu,r)\leftrightarrow(\mu,s)+(\nu,t)$.
+$(\mu,q)+(\nu,r)\leftrightarrow(\mu,s)+(\nu,t)$.  Note that
+\eqref{eq:H_int_start} and \eqref{eq:H_int_fields} neglect
+momentum-dependent corrections to $s$-wave interactions, which are
+generally captured (in a low-energy limit) by the use of an
+energy-dependent interaction pseudo-potential \cite{blume2002fermi}.
+These corrections will appear only at third order in our effective
+theory, as corrections to two-body interaction energies, and do not
+affect our discussions of multi-body interactions.  We therefore defer
+further discussion of momentum-dependent corrections to the end of
+section \ref{sec:two_body}, and their calculation to Appendix
+\ref{sec:momentum_dependence}.
 
@@
 such a gauge is guaranteed by the analytic properties of the Wannier
-orbitals $\phi_{in}$ \cite{kohn1959analytic}.
-
-We are interested in the ground and low-lying orbital eigenstates of
-several interacting atoms on a single lattice site, as such states can
-be readily prepared and coherently addressed in current experiments
-with ultracold atoms \cite{goban2018emergence}.  Although these
-experiments operate at temperatures sufficiently low for negligible
-thermal occupation of excited motional states, simply neglecting
-excited single-particle motional states fails to reproduce the
-observed spectra of multiply-occupied lattice sites.  The reason for
-this failure lies in the fact that the bare two-body Hamiltonian
-$H_{\t{int}}$ in \eqref{eq:H_int} is not diagonal in the basis of
-single-particle motional states.  Even at zero temperature, therefore,
-the presence of excited motional states modifies many-body atomic
-energy spectra when the magnitude of off-diagonal elements of
-$H_{\t{int}}$ approach the scale of low-lying single-particle motional
-state energy gaps.  To address this problem, we develop a low-energy
-effective theory for ultracold two-level fermionic atoms on a lattice.
-As our work is motivated by progress in experiments which initialize
-all atoms in identical (i.e.~ground) orbital and motional states, to
-simplify our theory we assume that any $N$ atoms on a single lattice
-site occupy $N$ nuclear spin states.  Multiple occupation of a single
-nuclear spin state on the same lattice site is initially forbidden by
-fermionic statistics, and subsequently cannot be violated in the
-absence of tunneling and hyperfine coupling.
+wavefunctions $\phi_{in}$ \cite{kohn1959analytic}.
+
+Current experiments with ${}^{87}$Sr can prepare up to five atoms
+atoms in the same (ground) orbital state on a single lattice site, and
+coherently address states with a single orbital excitation per lattice
+site (see Appendix \ref{sec:experiment_summary})
+\cite{goban2018emergence}.  The temperature of the atoms in these
+experiments is much smaller than single-atom motional excitation
+energies $\Delta_n\equiv E_n-E_0$ for $n>0$, which implies a
+negligible thermal occupation of excited motional states.  A naive
+calculation of the multi-body orbital excitation spectrum, however,
+assuming that atoms only occupy motional ground states with spatial
+wavefunctions $\phi_{i,0}$, fails to reproduce the experimentally
+observed orbital excitation spectrum of multiply-occupied lattice
+sites.  The reason that such a calculation fails to reproduce the
+observed excitation spectrum has to do with the fact that the two-body
+Hamiltonian $H_{\t{int}}$ in \eqref{eq:H_int} is not diagonal in the
+basis of single-particle motional eigenstates (i.e.~$\phi_{in}$).  For
+interacting atoms, the true motional ground states are therefore
+admixtures of the non-interacting motional eigenstates; when the
+off-diagonal elements of $H_{\t{int}}$ are comparable in magnitude to
+the excitation energies $\Delta_n$, this admixture can have
+non-negligible weight outside the non-interacting motional
+ground-state subspace.  In order to compute the correct many-body
+orbital excitation spectrum, we develop a low-energy effective theory
+for such atoms.  As our work is motivated by experiments that
+initialize all atoms in identical (i.e.~ground) orbital and motional
+states, to simplify our theory we assume that any $N$ atoms on a
+single lattice site occupy $N$ nuclear spin states.  Ground-state
+preparation implies that multiple occupation of a single nuclear spin
+state on any given lattice site is initially forbidden by fermionic
+statistics, and subsequently cannot occur in the absence of inter-site
+effects or hyperfine coupling.
 
@@
 \begin{align}
-  H_{\t{int}}^{\t{eff}} = \sum_{M=2}^{2I+1} H_M,
-\end{align}
-where $H_M$ is an $M$-body Hamiltonian, $I$ is the total nuclear spin
-of each atom (e.g.~$I=9/2$ for ${}^{87}$Sr), and the sum terminates at
+  H_{\t{int}}^{\t{eff}} = \sum_{M=2}^{2I+1} \sum_{p\ge1} H_M^{(p)},
+\end{align}
+where $H_M^{(p)}$ is an $M$-body Hamiltonian of order $p$ in the
+coupling constants $G^{qr}_{st}$, $I$ is the total nuclear spin of
+each atom (e.g.~$I=9/2$ for ${}^{87}$Sr), and the sum terminates at
 $2I+1$ because this is the largest number of atoms which may initially
-occupy a single lattice site.  Focusing on the low-lying orbital state
-excitations of this theory, we find that SU($N$) symmetry allows us to
-express multi-body Hamiltonians in the simple form
+occupy a single lattice site.  We explicitly compute all $M$-body
+Hamiltonians $H_M\equiv\sum_p H_M^{(p)}$ through order $p=3$, yielding
+effective two-, three-, and four-body interactions.  To facilitate
+understanding and comparison with experimental measurements of
+density-dependent atomic energy spectra [34], in which states with
+multiple orbital excitations on a single lattice site are
+spectroscopically off-resonant, we also consider a restriction of our
+theory to states with at most one orbital excitation per lattice site.
+In this case, we find that SU($N$) symmetry allows us to express
+multi-body Hamiltonians in the simple form
 \begin{align}
@@
 associated with each many-body internal atomic state.  We can then
-ignore all excited many-body motional states, as they will be neither
-thermally occupied nor externally interrogated.  This procedure makes
-the many-body problem more tractable by greatly reducing the
-dimensionality of our the relevant Hilbert space.
+ignore all excited motional states, as they will be neither thermally
+occupied nor externally interrogated.  This procedure makes the
+many-body problem more tractable by greatly reducing the
+dimensionality of the relevant Hilbert space.
 
@@
 where the blob on the right schematically represents the net effective
-two-body interaction.  In general, the sums over excited states in the
-loop diagrams of \eqref{eq:H_2_expansion} will diverge
-\cite{johnson2012effective}, which implies the need to renormalize the
-coupling constants in our theory.  Even if these diagrams do not
-diverge, we would like to express the effective two-body interactions
-in terms of net, physically observable two-body coupling constants as
-on the right-hand side of \eqref{eq:H_2_expansion}, rather than in
-terms of long sums at all orders of the bare coupling constants.  We
-therefore renormalize our coupling constants by introducing
+two-body interaction.  On physical grounds, the net effective two-body
+interaction must clearly be finite, but individual sums over excited
+states in the loop diagrams of \eqref{eq:H_2_expansion} may generally
+diverge \cite{johnson2012effective}.  These divergences ultimately
+appear due to our use of effective field theory to describe
+inter-atomic interactions in \eqref{eq:H_int_start}, rather than a
+detailed microscopic description of two-atom scattering.  Divergences
+of this sort are a generic feature of field theories, and can be dealt
+with using standard techniques, in particular through renormalization.
+We therefore renormalize our coupling constants by introducing
 counter-terms $\tilde G_X$ into the effective interaction Hamiltonians
-via $G_X\to G_X+\tilde G_X$.  The values of these counter-terms are
-fixed by enforcing that $G_X$ is equal to the net two-body interaction
-strength.  Diagrammatically, this renormalization condition is
+via $G_X\to G_X+\tilde G_X$.  The values of counter-terms can be fixed
+by enforcing that $G_X$ is equal to the net two-body interaction
+strength, which has the added benefit of allowing us to express the
+effective two-body interactions in terms of net, physically observable
+two-body coupling constants on the right-hand side of
+\eqref{eq:H_2_expansion}, rather than in terms of long sums at all
+orders of the bare coupling constants.  Diagrammatically, this
+renormalization condition is
 \begin{align}
@@
 where we represent counter-terms by a crossed dot (i.e.~$\otimes$) at
-a vertex. To leading order in the coupling constants, it follows that
-the counter-terms are second order in the coupling constants,
-i.e.~$\tilde G\sim G^2$.  This finding implies that counter-terms only
-come into play at next-to-leading order (NLO) in the calculation of
-any effective $M$-body interaction Hamiltonian.  By construction,
-these counter-terms exactly cancel all terms beyond leading order in
+a vertex. To leading order, it follows that the counter-terms are
+second order in the coupling constants, i.e.~$\tilde G\sim G^2$.  This
+finding implies that counter-terms only come into play at
+next-to-leading order (NLO) in the calculation of any effective
+$M$-body interaction Hamiltonian.  By construction, these
+counter-terms exactly cancel all terms beyond leading order in
 \eqref{eq:H_2_expansion}, which implies that the effective two-body
@@
 consequently our result in \eqref{eq:H_2}, neglects the effect of
-momentum-dependent two-body scattering.  When the effective range of
-momentum-dependent interactions is comparable to the scattering
-lengths $a_X$, as is the case for ultracold ${}^{87}$Sr, these
-interactions are third order in the coupling constants $G_X$.  Just as
-the $\O\p{G^2}$ counter-terms do not affect $M$-body interactions
-until next-to-leading order (NLO), the $\O\p{G^3}$ momentum-dependent
-interactions do not come into play until next-next-leading order
-(NNLO).  Given that we develop our low-energy effective theory through
-third order in the coupling constants, these interactions will not
-appear in any of our three- and four-body calculations, but they do
-need to be considered in the calculation of pair-wise interaction
-energies.  The primary interest of our work, however, concerns
-effective $M$-body interactions for $M\ge3$; we therefore defer the
-calculation of momentum-dependent two-body interactions to Appendix
-\ref{sec:momentum_dependence}.
+momentum-dependent two-body scattering \cite{blume2002fermi}.  When
+the effective range of momentum-dependent interactions is comparable
+to the scattering lengths $a_X$, as is the case for ultracold
+${}^{87}$Sr, these interactions are third order in the coupling
+constants $G_X$.  Just as the $\O\p{G^2}$ counter-terms do not affect
+$M$-body interactions until next-to-leading order (NLO), the
+$\O\p{G^3}$ momentum-dependent interactions do not come into play
+until next-next-leading order (NNLO).  Given that we develop our
+low-energy effective theory through third order in the coupling
+constants, these interactions will not appear in any of our three- and
+four-body calculations, but they do need to be considered in the
+calculation of pair-wise interaction energies.  The primary interest
+of our work, however, concerns effective $M$-body interactions for
+$M\ge3$; we therefore defer the calculation of momentum-dependent
+two-body interactions to Appendix \ref{sec:momentum_dependence}.
 
@@
 excitations for up to five atoms per lattice site
-\cite{goban2018emergence}.  Due to the SU($N$) symmetry of
-inter-atomic interactions, manifest in the fact that all coupling
-constants are independent of nuclear spin, we can write all effective
-$M$-body Hamiltonians which address these states in the form
+\cite{goban2018emergence} (see Appendix \ref{sec:experiment_summary}
+for a summary of relevant experimental procedures).  Due to the
+SU($N$) symmetry of inter-atomic interactions, manifest in the fact
+that all coupling constants are independent of nuclear spin, we can
+write all effective $M$-body Hamiltonians which address these states
+in the form
 \begin{align}
@@
   = \sum_X E_X \P_X
-  + \sum_\mu\p{\Delta T_\mu^z + \Omega_\mu T_\mu^x},
+  + \sum_\mu\p{\Delta T_\mu^\z + \Omega_\mu T_\mu^\x},
   \label{eq:H_rabi}
@@
 \begin{align}
-  T_\mu^z \equiv \f12 \p{\c_{\mu,\e}^\dag \c_{\mu,\e}
+  T_\mu^\z \equiv \f12 \p{\c_{\mu,\e}^\dag \c_{\mu,\e}
     - \c_{\mu,\g}^\dag \c_{\mu,\g}},
   &&
-  T_\mu^x \equiv \f12 \p{\c_{\mu,\e}^\dag \c_{\mu,\g}
+  T_\mu^\x \equiv \f12 \p{\c_{\mu,\e}^\dag \c_{\mu,\g}
     + \c_{\mu,\g}^\dag \c_{\mu,\e}},
@@
 \begin{align}
-  H_{\N X} = \delta S_{\N X}^z + \Omega_{\N X} S_{\N X}^x,
+  H_{\N X} = \delta S_{\N X}^\z + \Omega_{\N X} S_{\N X}^\x,
   \label{eq:H_NX}
@@
 \begin{align}
-  S_{\N X}^z &\equiv \f12\p{\op{\N X}-\op{\N,0}},
+  S_{\N X}^\z &\equiv \f12\p{\op{\N X}-\op{\N,0}},
   \label{eq:S_NX_z} \\
-  S_{\N X}^x &\equiv \f12\p{\op{\N X}{\N,0}+\op{\N,0}{\N X}},
+  S_{\N X}^\x &\equiv \f12\p{\op{\N X}{\N,0}+\op{\N,0}{\N X}},
   \label{eq:S_NX_x}
@@
     mixture of nuclear spins.  Atoms are prepared in the ground state
-    of a lattice with depth $\U=54E_R$, where
-    $E_R\approx3.5\times2\pi~\t{kHz}$ is the lattice photon recoil
-    energy of the atoms, and interrogated by a laser with Rabi
+    of a lattice with depth $\U=54 E_{\t{R}}$, where
+    $E_{\t{R}}\approx3.5\times2\pi~\t{kHz}$ is the lattice photon
+    recoil energy of the atoms, and interrogated by a laser with Rabi
     frequency $\Omega_I=50\times2\pi~\t{Hz}$ for a time
@@
 states of a rectangular lattice with depths
-$\U=\p{\U_x,\U_y,\U_z}=\p{41,55,69}E_R$, where
-$E_R\approx3.5\times2\pi~\t{kHz}$ is the lattice photon recoil energy
-of the atoms.  Such samples can be prepared in experiments which can
-vary the single-site occupation number $N$, and which can control for
-the total number of atoms that are addressed by an external
-interrogation laser.  Figure \ref{fig:sweep} shows the population of
-the excited ${}^3P_0$ orbital state when these atoms are interrogated
-for a time $t=\pi/\Omega_I$ by a laser with Rabi frequency
-$\Omega_I=50\times2\pi~\t{Hz}$ (i.e.~for individual atoms with nuclear
-spin $\mu=I$) and detuning $\Delta$ from the single-atom
+$\U=\p{\U_\x,\U_\y,\U_\z}=\p{41,55,69} E_{\t{R}}$, where
+$E_{\t{R}}\approx3.5\times2\pi~\t{kHz}$ is the lattice photon recoil
+energy of the atoms.  Such samples can be prepared in experiments
+which can vary the single-site occupation number $N$, and which can
+control for the total number of atoms that are addressed by an
+external interrogation laser.  Figure \ref{fig:sweep} shows the
+population of the excited ${}^3P_0$ orbital state when these atoms are
+interrogated for a time $t=\pi/\Omega_I$ by a laser with Rabi
+frequency $\Omega_I=50\times2\pi~\t{Hz}$ (i.e.~for individual atoms
+with nuclear spin $\mu=I$) and detuning $\Delta$ from the single-atom
 ${}^1S_0\to{}^3P_0$ orbital excitation energy.  The ${}^3P_0$
@@
   \rho_N^{(\pm)}\p{t} = \f1{\Tr\P_{N,0}}
-  \sum_{\abs{\N}=N} \exp\p{-it\Omega_{\N,\pm}S_{\N,\pm}^x}
-  \op{\N,0} \exp\p{it\Omega_{\N,\pm}S_{\N,\pm}^x},
+  \sum_{\abs{\N}=N} \exp\p{-it\Omega_{\N,\pm}S_{\N,\pm}^\x}
+  \op{\N,0} \exp\p{it\Omega_{\N,\pm}S_{\N,\pm}^\x},
 \end{align}
 where the Rabi frequencies $\Omega_{\N,+},\Omega_{\N,-}$ and
-pseudo-spin operators $S_{\N,\pm}^x$ are respectively given in
+pseudo-spin operators $S_{\N,\pm}^\x$ are respectively given in
 \eqref{eq:O_N+}, \eqref{eq:O_N-}, and \eqref{eq:S_NX_x}.  Denoting the
-eigenstates of $S_{\N,\pm}^x$ by
+eigenstates of $S_{\N,\pm}^\x$ by
 \begin{align}
@@
 emergent multi-body interactions that inherit the SU($N$) symmetry of
-the atoms' bare, two-body interactions.  The SU($N$) symmetry of
-$M$-body Hamiltonians in the effective theory has allowed us to
-express them in a simple form, and fully characterize their
-corresponding eigenstates and spectra.  Capitalizing on the extreme
-precision of state-of-the-art clock spectroscopy, we have tested
-spectral predictions of our theory against direct experimental
-measurements of the many-body ${}^{87}$Sr excitation spectrum.  This
-comparison shows good agreement between theory and experiment, clearly
-demonstrating the need to consider multi-body effects for
-understanding the low-energy physics of high density AEA samples on a
-3-D lattice.  Finally, we analyzed the many-body orbital-state
-dynamics of multiply-occupied lattice sites prepared in a nuclear spin
-mixture and interrogated via Rabi spectroscopy.  This analysis is
-useful for future experimental probes of many-body state structures,
-as well as for the preparation of long-lived states with multi-partite
-entanglement (i.e.~$\ket{\N,\pm}$) which may be used as a resource to
-perform quantum information processing tasks.
+the atoms' bare, two-body interactions.  When considering the
+low-lying orbital excitations of our effective theory, the SU($N$)
+symmetry of $M$-body Hamiltonians allowed us to express them in a
+simple form, and to fully characterize their eigenstates and spectra.
+Capitalizing on the extreme precision of state-of-the-art clock
+spectroscopy, we have tested spectral predictions of our theory
+against direct experimental measurements of the many-body ${}^{87}$Sr
+excitation spectrum.  This comparison shows good agreement between
+theory and experiment, clearly demonstrating the need to consider
+multi-body effects for understanding the low-energy physics of high
+density AEA samples on a 3-D lattice.  Finally, we analyzed the
+many-body orbital-state dynamics of multiply-occupied lattice sites
+prepared in a nuclear spin mixture and interrogated via Rabi
+spectroscopy.  This analysis is useful for future experimental probes
+of many-body state structures, as well as for the preparation of
+long-lived states with multi-partite entanglement
+(i.e.~$\ket{\N,\pm}$) which may be used as a resource to perform
+quantum information processing tasks.
 
@@
 improvement in the form of controlled, systematic treatment of
-higher-order and tunneling processes.
-% These efforts will be necessary to go beyond the deep-lattice limit
-% and study inter-site orbital-state and nuclear spin exchange
-% dynamics.  Nonetheless, our work makes a major step towards the
-% experimental investigation of multi-body SU($N$) physics.
-Nonetheless, our work makes a major step towards the experimental
-investigation of multi-body SU($N$) physics, providing the necessary
-framework for future studies going beyond the deep-lattice limit to
-realize multi-body super-exchange dynamics and orbital SU($N$) quantum
-magnetism with AEAs.
+higher-order and tunneling processes.  Nonetheless, our work makes a
+major step towards the experimental investigation of multi-body
+SU($N$) physics, providing the necessary framework for future studies
+going beyond the deep-lattice limit to realize multi-body
+super-exchange dynamics and orbital SU($N$) quantum magnetism with
+AEAs.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Summary of experimental procedures}
+\label{sec:experiment_summary}
+
+The work in this paper is closely tied to the experimental work in
+ref.~\cite{goban2018emergence}; for quick reference, we here provide a
+short summary of the experimental procedures used therein.  The
+experiment begins by preparing a degenerate gas of $10^4$-$10^5$
+(fermionic) ${}^{87}$Sr atoms, in a uniform mixture of their ten
+nuclear spin states, at $\sim0.1$ of their fermi temperature (10 to 20
+nanokelvin) \cite{campbell2017fermidegenerate, marti2018imaging}.
+This gas is loaded into a primitive cubic optical lattice at the
+``magic wavelength'' for which both ground (${}^1S_0$) and
+first-excited (${}^3P_0$) electronic orbital states of the atoms
+experience the same lattice potential.  Lattice depths along the
+principal axes of the lattice are comparable in magnitude, with a
+geometric mean that can be varied from 30 to 80 $E_{\t{R}}$, where
+$E_{\t{R}}\approx3.5\times2\pi~\t{kHz}$ is the lattice photon recoil
+energy of the atoms.  These lattice depths are sufficiently large as
+to neglect tunneling on the time scales relevant to the experiment.
+Atoms are then addressed by an external (``clock'') interrogation
+laser with an ultranarrow (26 mHz) linewidth, detuned by $\Delta$ from
+the single-atom ${}^1S_0-{}^3P_0$ transition frequency.  When
+$\Delta=0$, non-interacting atoms in singly-occupied lattice sites
+thus undergo Rabi oscillations.  After a chosen time, the experiment
+removes all ground-state (${}^1S_0$) atoms from the lattice and uses
+absorption imaging to identify the remaining excited-state (${}^3P_0$)
+atoms.  As discussed in section \ref{sec:spectra}, a similar procedure
+with $\Delta\ne0$ leads to the spectrum shown in figure
+\ref{fig:sweep}, thereby constituting a measurement of the many-body
+atomic excitation spectrum.  The many-body atomic excitation spectrum
+is different from the single-body atomic excitation spectrum due to
+the presence of inter-atomic interactions.  We note that this
+spectroscopic protocol addresses only singly-excited orbital states of
+any lattice site.  Doubly-excited states are manifestly off-resonant
+for single-photon absorption processes, while multi-photon absorption
+processes are always off-resonant due to
+\begin{enumerate*}
+\item the non-linearity (on the scale of kHz) of on-site electronic
+  orbital excitation energies (resulting from non-uniform coupling
+  constants $G_X$), combined with
+\item the ultranarrow linewidth (on the scale of mHz) of the
+  interrogation laser.
+\end{enumerate*}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Derivation of the effective Hamiltonian expansion}
@@
 operator $V$ which weakly couples $\G_0$ and $\E_0$.  For all
-$\ket{\psi_g}\in\G_0$ and $\ket{\psi_e},\ket{\chi_e}\in\E_0$, we have
-$\bk{\psi_g|H_0|\psi_g}=0$, $\bk{\psi_e|H_0|\psi_g}=0$, and
-$\abs{\bk{\psi_e|V|\psi_g}}\ll\bk{\chi_e|H_0|\chi_e}$.  The net
+$\ket{\psi_\g}\in\G_0$ and $\ket{\phi_\e},\ket{\chi_\e}\in\E_0$, we
+have $\bk{\psi_\g|H_0|\psi_\g}=0$, $\bk{\phi_\e|H_0|\psi_\g}=0$, and
+$\abs{\bk{\phi_\e|V|\psi_\g}}\ll\bk{\chi_\e|H_0|\chi_\e}$.  The net
 Hamiltonian $H=H_0+V$ will naturally admit a decomposition of the
@@
 $X\in\set{\g\g,\e\g^-,\e\g^+,\e\e}$ on coupling constants $G_X$ in the
-remainder of this appendix.  To further simplify notation, we will
+remainder of this Appendix.  To further simplify notation, we will
 also neglect the explicit dependence of parameters on the lattice
@@
 \begin{align}
-  \U\sin^2\p{k_L\cdot x} \approx \U \p{k_L^x k_L^y k_L^z}^{2/3} x^2
-  = \f12 m_A \omega_{\t{eff}}^2 x^2,
-  &&
-  \omega_{\t{eff}} \equiv \sqrt{2~ \U k_L^2 / m_A},
-\end{align}
-where $k_L=\p{k_L^x,k_L^y,k_L^z}$ is the lattice wavenumber, $m_A$ is
-the atomic mass, and $\omega_{\t{eff}}$ is an effective angular
-harmonic trap frequency.  We then use $\omega_{\t{eff}}$ in place of
-$\omega$ in \eqref{eq:harmonic_gamma}, and use an overlap integral $K$
-computed with the ground-state wavefunctions $\phi_0$ in a lattice
-rather than those in a harmonic oscillator.  We retrieve free-space
-$s$-wave scattering lengths $a_{\t{free}}$ for ${}^{87}$Sr from
-ref.~\cite{zhang2014spectroscopic} to determine the free-space
-coupling constants $G_{\t{free}} \equiv \p{4\pi/m_A} a_{\t{free}}$.
-This procedure yields an effective coupling constant $G_{\t{lattice}}$
-given by
+  \U\sin^2\p{k_{\t{L}} \cdot x}
+  \approx \U \sp{\p{k_{\t{L}}^\x k_{\t{L}}^\y k_{\t{L}}^\z}^{1/3} x}^2
+  = \f12 \mA \omega_{\t{eff}}^2 x^2,
+  &&
+  \omega_{\t{eff}} \equiv \sqrt{2~ \U k_{\t{L}}^2 / \mA},
+\end{align}
+where $k_{\t{L}}=\p{k_{\t{L}}^\x,k_{\t{L}}^\y,k_{\t{L}}^\z}$ is the
+lattice wavenumber, $\mA$ is the atomic mass, and $\omega_{\t{eff}}$
+is an effective angular harmonic trap frequency.  We then use
+$\omega_{\t{eff}}$ in place of $\omega$ in \eqref{eq:harmonic_gamma},
+and use an overlap integral $K$ computed with the ground-state
+wavefunctions $\phi_0$ in a lattice rather than those in a harmonic
+oscillator.  We retrieve free-space $s$-wave scattering lengths
+$a_{\t{free}}$ for ${}^{87}$Sr from ref.~\cite{zhang2014spectroscopic}
+to determine the free-space coupling constants
+$G_{\t{free}} \equiv \p{4\pi/\mA} a_{\t{free}}$.  This procedure
+yields an effective coupling constant $G_{\t{lattice}}$ given by
 \begin{align}
@@
   &&
-  \chi \equiv \sqrt{2}~ \xi~ \f{\p{m_AC_6}^{1/4}}{a},
+  \chi \equiv \sqrt{2}~ \xi~ \f{\p{\mA C_6}^{1/4}}{a},
   \label{eq:effective_range}
@@
 \eqref{eq:coupling_tensor}, but with scattering lengths
-$a\to r_{\t{eff}}a^2/2$ and the effective range defined by
-\eqref{eq:effective_range} for each scattering length with the
+$a\to r_{\t{eff}}a^2/2$ and the effective range $r_{\t{eff}}$ defined
+by \eqref{eq:effective_range} for each scattering length with an
 appropriate $C_6$ coefficient.  We retrieve $C_6$ coefficients for
@@
 single- and two-body Hamiltonians: we assumed that
-\begin{enumerate*}[label=\roman*)]
+\begin{enumerate*}
 \item tunneling between lattice sites and
@@
 where the coefficients can be expanded as
-$U_{M,X} = \sum_p U_{M,X}^{(p)}$ with terms $U_{M,X}^{(p)}$ at order
-$p$ in the coupling constants $G_Y$.  The terms $U_{M,X}^{(p)}$ can be
+$U_{MX} = \sum_p U_{MX}^{(p)}$ with terms $U_{MX}^{(p)}$ at order $p$
+in the coupling constants $G_Y$.  The terms $U_{MX}^{(p)}$ can be
 determined from the $M$-body $p$-order Hamiltonians $H_M^{(p)}$

### 2019-02-02 soc_squeezing/paper/collective_spin.tex
fixed typo with normal vs. mathcal S symbol

@@
 general product
-$\S_{\v n}\S_{\v m}=\sum_{\v\ell}f_{\v n\v m\v\ell}S_{\v\ell}$ worked
+$\S_{\v n}\S_{\v m}=\sum_{\v\ell}f_{\v n\v m\v\ell}\S_{\v\ell}$ worked
 out in Appendices \ref{sec:identities}--\ref{sec:prod_general}, while

### 2019-02-04 multi_body_interactions/papers/theory/reviews_NJP/referee_1_response.tex
first batch of responses to referee 1 for NJP

\documentclass[preprint]{revtex4-1}

% linking references
\usepackage{hyperref}
\hypersetup{
  breaklinks=true,
  colorlinks=true,
  linkcolor=blue,
  urlcolor=cyan,
}

% general physics / math packages and commands
\usepackage{physics,amsmath,amssymb,braket,dsfont,bm}
\renewcommand{\t}{\text} % text in math mode
\newcommand{\f}{\dfrac} % shorthand for fractions
\newcommand{\p}[1]{\left(#1\right)} % parenthesis
\renewcommand{\sp}[1]{\left[#1\right]} % square parenthesis
\renewcommand{\set}[1]{\left\{#1\right\}} % curly parenthesis
\newcommand{\bk}{\braket} % shorthand for braket

\newcommand{\g}{\text{g}}
\newcommand{\e}{\text{e}}
\newcommand{\x}{\text{x}}
\newcommand{\y}{\text{y}}
\newcommand{\z}{\text{z}}

\renewcommand{\c}{\hat{c}}
\newcommand{\n}{\hat{n}}

\newcommand{\A}{\mathcal{A}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\G}{\mathcal{G}}
\renewcommand{\H}{\mathcal{H}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\K}{\mathcal{K}}
\renewcommand{\L}{\mathcal{L}}
\newcommand{\M}{\mathcal{M}}
\newcommand{\N}{\mathcal{N}}
\renewcommand{\O}{\mathcal{O}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\Q}{\mathcal{Q}}
\renewcommand{\S}{\mathcal{S}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\1}{\mathds{1}}

% inline lists
\usepackage[inline]{enumitem}
\setlist[enumerate]{leftmargin=*} % nice margins for enumerate

% color definitions (used in a figure)
\usepackage{xcolor}
\newcommand{\blue}[1]{\textcolor{blue}{#1}}
\newcommand{\red}[1]{\textcolor{red}{#1}}
\newcommand{\green}[1]{\textcolor{green}{#1}}
\definecolor{lightblue}{RGB}{31,119,180}
\definecolor{orange}{RGB}{255,127,14}
\definecolor{green}{RGB}{44,160,44}
\definecolor{lightred}{RGB}{214,39,40}

% for strikeout text
% normalem included to prevent underlining titles in the bibliography
\usepackage[normalem]{ulem}

\begin{document}

\section*{Referee 1}

We thank the referee for taking the time to carefully read our
manuscript, and for providing thoughtful and constructive feedback.
Below, we address the comments and concerns brought up by the referee.
We hope that the referee finds our responses and revisions
satisfactory, deeming our manuscript acceptable for publication in New
Journal of Physics.

Excerpts from the referee reports are written in \blue{blue}, excerpts
from our original manuscript are written in \red{red}, and new text in
the revised version of our manuscript is written in \green{green}.
All page, line, and reference numbers generally refer to those in our
original manuscript.

\begin{enumerate}
\item Concerning:

  \blue{The authors study the effective on-site interactions for the
    description of several fermionic particles confined in a very deep
    optical lattice. The focus is on alkaline earth atoms, where the
    s-wave interaction strength is independent on the nuclear spin,
    and on all atoms in the lowest state of the trapping
    potential. The manuscript is an extended version on the
    theoretical tools applied in the comparison between experiment and
    theory of Ref. [34].  The analysis is very important and
    timely. However, in the present form, the manuscript has some
    significant short comings, which have to be improved before a
    decision about publication can be made.}

  \blue{One of the main points of criticism is that the manuscript is
    very poorly written and very difficult to access. A few examples
    are listed below:}

  We thank the referee for the succinct summary of our work, and for
  recognizing that our analysis is ``very important and timely''.  We
  hope that thew revisions outlined below have made our manuscript
  more accessible, and that we have addressed shortcomings identified
  by the reviewer to a satisfactory degree.

\item Concerning the comment:

  \blue{(i) The authors try to be rather general in their description
    such as title, abstract and introduction. During the description
    of the results, they start to apply several approximations, which
    are neither mentioned before or justified.  As an example: the
    authors claim to derive a general low energy theory for orbital
    and nuclear spin stats. However, in Eq.(10) the authors suddenly
    restrict the analysis to a low density of excitations in the ``e''
    orbital state. This restriction is never explained nor motivated.}

  \label{pt:general}

  The referee is correct that we do not ``derive a general low energy
  theory for orbital and nuclear spin states''.  In our abstract, for
  example, we merely claim that (page 1, line 27)

  \red{Motivated in particular by recent experiments ... we develop a
    low-energy effective theory ... which exhibits emergent multi-body
    SU($N$)-symmetric interactions ...}

  In our introduction, meanwhile, the very first sentence discussing
  the technical contributions of our manuscript (page 4, line 19)
  explicitly states that

  \red{In this work, we investigate the first experimental
    capabilities with ultracold fermionic AEAs to prepare high-density
    samples in a 3-D lattice with multiple occupation of individual
    lattice sites [34].}

  We similarly begin both our abstract and introduction with a
  discussion of specific experiments, and make frequent reference to
  experimental developments and capabilities.  We thus believe that
  our presentation is faithful to the technical content of our
  manuscript; we do not make any claim of having developed a
  completely general theory two-orbital multi-body SU($N$)-symmetric
  interactions.  We claim only to have developed a low-energy theory
  that (i) is relevant for current experiments, and (ii) exhibits
  multi-body SU($N$)-symmetric interactions.  For these reasons, we do
  indeed introduce several approximations even in the first paragraph
  of our theory overview (section II, page 5).  We believe that
  theoretical work motivated by specific experiments is generally
  justified in making approximations that are relevant to the
  parameter regimes of these experiments.

  Nonetheless, we fully recognize and agree with the referee on the
  importance of making the scope of our technical contributions clear
  (as the referee comments later in point \ref{pt:restrictions}).  The
  fact that the referee at any point believed we will present a
  general theory without applying several experimentally-motivated
  approximations clearly signifies that we need to be more explicit in
  our aims.  Consequently, we have modified the second sentence of our
  original abstract, from (page 1, line 27)

  \red{Motivated in particular by recent experiments ...}

  to

  \green{In order to analyze recent experiments ...}

  We have also clarified an important point in our introduction,
  replacing the following text (page 4, line 22)

  \red{Specifically, we consider the preparation of isolated few-body
    systems in the deep-lattice limit, ...}

  with

  \green{Specifically, we consider ground-state preparation of
    isolated few-body systems in the deep-lattice limit, ...}

  The point on ground-state preparation in the deep lattice limit is
  important for a restriction we make later concerning single
  occupation of nuclear spin states on individual lattice sites.  This
  restriction (discussed in point \ref{pt:single_nuclear} below)
  together with the experimentally-motivated approximations made in
  the first paragraph of the theory overview (section II, page 5)
  constitute all key restrictions of our theory.

  We hope that these changes, together with those discussed in the
  examples below (points \ref{pt:orbital_state} and
  \ref{pt:single_nuclear}), clarify the motivation and scope of our
  manuscript, such that a reader is not led to believe that we present
  a general theory without applying several well-motivated
  approximations that are relevant to current experiments.

\item Concerning the specific example:

  \blue{However, in Eq.(10) the authors suddenly restrict the analysis
    to a low density of excitations in the ``e'' orbital state. This
    restriction is never explained nor motivated.}

  \label{pt:orbital_state}

  It is true, as the referee points out, that in Eq.~(10) (page 8) we
  present an effective Hamiltonian addressing only states with at most
  one orbital excitation (``e'') per lattice site.  We motivate this
  restriction immediately after Eq.~(8) in our theory overview, as
  soon as we begin a technical discussion of multi-body interactions
  (page 7, line 20):

  \red{We are interested in the ground and low-lying orbital
    eigenstates of several interacting atoms on a single lattice site,
    as such states can be readily prepared and coherently addressed in
    current experiments with ultracold atoms [34].}

  We acknowledge that the referee found the paragraph containing this
  sentence confusing, and have consequently rewritten the paragraph
  between Eqs.~(8) and (9) entirely (see point \ref{pt:confusion}).
  In any case, our actual results in section III (page 8) take the
  general form of Eq.~(9) (page 7), and make no restrictions on the
  number of orbital excitations per lattice site.  The expression in
  Eq.~(10) is merely a simplified restriction of our technical
  results, made (i) to highlight important features of the low-lying
  orbital excitations in our multi-body SU($N$) theory, and (ii) to
  facilitate comparison with optical lattice clock experiments.

  To clarify this point, we have made more explicit the fact that our
  actual results take the form of Eq.~(9), and added additional
  discussion of Eq.~(10).  Specifically, we have replaced the
  following text (page 7, line 48):

  \red{Our low-energy effective theory exhibits SU($N$)-symmetric
    multi-body interactions, such that the effective interaction
    Hamiltonian can be written in the form
    \begin{align}
      H_{\t{int}}^{\t{eff}} = \sum_{M=2}^{2I+1} H_M, \tag{9}
    \end{align}
    where $H_M$ is an $M$-body Hamiltonian, $I$ is the total nuclear
    spin of each atom (e.g.~$I=9/2$ for ${}^{87}$Sr), and the sum
    terminates at $2I+1$ because this is the largest number of atoms
    which may initially occupy a single lattice site.  Focusing on the
    low-lying orbital state excitations of this theory, we find that
    SU($N$) symmetry allows us to express multi-body Hamiltonians in
    the simple form}

  with

  \green{Our low-energy effective theory exhibits SU($N$)-symmetric
    multi-body interactions, such that the effective interaction
    Hamiltonian can be written in the form
    \begin{align}
      H_{\t{int}}^{\t{eff}} = \sum_{M=2}^{2I+1} \sum_{p\ge1} H_M^{(p)},
      \tag{9}
    \end{align}
    where $H_M^{(p)}$ is an $M$-body Hamiltonian of order $p$ in the
    coupling constants $G^{qr}_{st}$, $I$ is the total nuclear spin of
    each atom (e.g.~$I=9/2$ for ${}^{87}$Sr), and the sum terminates
    at $2I+1$ because this is the largest number of atoms which may
    initially occupy a single lattice site.  We explicitly compute all
    $M$-body Hamiltonians $H_M\equiv\sum_p H_M^{(p)}$ through order
    $p=3$, yielding effective two-, three-, and four-body
    interactions.  To facilitate understanding and comparison with
    experimental measurements of density-dependent atomic energy
    spectra [34], in which states with multiple orbital excitations on
    a single lattice site are spectroscopically off-resonant, we also
    consider a restriction of our theory to states with at most one
    orbital excitation per lattice site.  In this case, we find that
    SU($N$) symmetry allows us to express multi-body Hamiltonians in
    the simple form}

  We have also modified the following sentence in our summary (page
  28, line 43):

  \red{The SU($N$) symmetry of $M$-body Hamiltonians allowed us to
    express them in a simple form, and fully characterize their
    corresponding eigenstates and spectra.}

  to:

  \green{When considering the low-lying orbitalexcitations of our
    effective theory, the SU($N$) symmetry of $M$-body Hamiltonians
    allowed us to express them in a simple form, and to fully
    characterize their eigenstates and spectra.}

  Finally, the new appendix we have added as a part of point
  \ref{pt:description} (below) contains additional discussion of the
  restriction to at most one orbital excitation per lattice site.

  We hope that these changes clarify the scope of our work, in
  particular highlighting that the theory developed in section III
  applies to any number of orbital excitations, despite the
  restriction to the single-excitation case for the sake of comparison
  with experiments in section IV (page 18).

\item Concerning the next example:

  \blue{In addition, the restriction to a single occupation of the
    nuclear spin state is never motivated or justified but just
    introduced at some point.}

  \label{pt:single_nuclear}

  The referee is correct that we restrict our theory to the case of
  single occupation of nuclear spin states on any given lattice site.
  We do so because (i) this regime is precisely that relevant to
  current experiments with ultracold atoms, which begin with
  ground-state preparation of the atoms, and (ii) this restriction
  greatly simplifies our theory.  We explained both the motivation and
  the justification for this restriction in the overview of our theory
  (page 7, line 40):

  \red{As our work is motivated by progress in experiments which
    initialize all atoms in identical (i.e.~ground) orbital and
    motional states, to simplify our theory we assume that any $N$
    atoms on a single lattice site occupy $N$ nuclear spin states.
    Multiple occupation of a single nuclear spin state on the same
    lattice site is initially forbidden by fermionic statistics, and
    subsequently cannot be violated in the absence of tunneling and
    hyperfine coupling.}

  Deferring the referee's later comment that this text was confusing
  to point \ref{pt:confusion}, we emphasize the fact that our theory
  is the first to include hyperfine structure when considering
  multi-body interactions of ultracold atoms on a lattice.  Of course,
  a more general theory relaxing the above restriction (i.e.~on the
  occupation of nuclear spin states) will eventually be necessary for
  future experiments with a shallower lattice or additional control
  fields.  Such a general theory will need to consider inter-site
  effects or hyperfine coupling (or, less likely, non-ground-state
  preparation protocols), as these are the mechanisms by which a
  nuclear spin state can become multiply-occupied on a single lattice
  site.  Given the that these mechanisms (i) are largely irrelevant
  for the experiments that prompted this work, and (ii) introduce
  major complications into the study of SU($N$) multi-body
  interactions, we believe that it is appropriate to leave such
  considerations for future work.

  For clarification, we have modified the sentence on page 7, line 43
  from

  \red{Multiple occupation of a single nuclear spin state on the same
    lattice site is initially forbidden by fermionic statistics, and
    subsequently cannot be violated in the absence of tunneling and
    hyperfine coupling.}

  to

  \green{Ground-state preparation implies that multiple occupation of
    a single nuclear spin state on any given lattice site is initially
    forbidden by fermionic statistics, and subsequently cannot occur
    in the absence of inter-site effects or hyperfine coupling.}

  We hope that this modification, together with the modifications
  mentioned in point \ref{pt:general}, clarifies our justification for
  restricting to the case of single occupation of nuclear spin states.

\item Concerning the following comments:

  \blue{The referee understands, that these restrictions are justified
    for the comparison with the experiment in Ref.[34]. However, it
    should be clearly pointed, that the analysis is not general and
    adapted to the description to these experiments in the abstract
    and introduction.}

  \label{pt:restrictions}

  We thank the referee for pointing out the potential to mislead
  readers on the generality of our theory, and hope that the new
  version of our manuscript does a better job of calibrating the
  expectations of a reader with regard to restrictions and
  approximations that we make.  In particular, we hope that the
  current text makes clear to readers that we are developing a theory
  for the purpose of understanding current optical lattice clock
  experiments, and not to understand SU($N$) multi-body interactions
  in full generality.

\item Concerning the description of the experiment with which we
  compare:

  \blue{Furthermore, a short description of the experiment would be
    mandatory for an independent manuscript.}

  \label{pt:description}

  We thank the referee for pointing out that a description of relevant
  experimental procedures would make our manuscript more
  self-contained, and we recognize the value in doing so.  We have
  therefore added a new appendix to our manuscript, summarizing all
  relevant experimental procedures:

  \green{{\bf Appendix A:~Summary of experimental procedures}}

  \green{The work in this paper is closely tied to the experimental
    work in ref.~[34]; for quick reference, we here provide a short
    summary of the experimental procedures used therein.  The
    experiment begins by preparing a degenerate gas of $10^4$-$10^5$
    (fermionic) ${}^{87}$Sr atoms, in a uniform mixture of their ten
    nuclear spin states, at $\sim0.1$ of their fermi temperature (10
    to 20 nanokelvin) [8, 52].  This gas is loaded into a primitive
    cubic optical lattice at the ``magic wavelength'' for which both
    ground (${}^1S_0$) and first-excited (${}^3P_0$) electronic
    orbital states of the atoms experience the same lattice potential.
    Lattice depths along the principal axes of the lattice are
    comparable in magnitude, with a geometric mean that can be varied
    from 30 to 80 $E_{\t{R}}$, where
    $E_{\t{R}}\approx3.5\times2\pi~\t{kHz}$ is the lattice photon
    recoil energy of the atoms.  These lattice depths are sufficiently
    large as to neglect tunneling on the time scales relevant to the
    experiment.  Atoms are then addressed by an external (``clock'')
    laser with an ultranarrow (26 mHz) linewidth, detuned by $\Delta$
    from the single-atom ${}^1S_0-{}^3P_0$ transition frequency.  When
    $\Delta=0$, non-interacting atoms in singly-occupied lattice sites
    thus undergo Rabi oscillations.  After a chosen time, the
    experiment removes all ground-state (${}^1S_0$) atoms from the
    lattice and uses absorption imaging to identify the remaining
    excited-state (${}^3P_0$) atoms.  As discussed in section IV, a
    similar procedure with $\Delta\ne0$ leads to the spectrum shown in
    figure 3, thereby constituting a measurement of the many-body
    atomic excitation spectrum.  The many-body atomic excitation
    spectrum is different from the single-body atomic excitation
    spectrum due to the presence of inter-atomic interactions.  We
    note that this spectroscopic protocol addresses only
    singly-excited orbital states of any lattice site.  Doubly-excited
    states are manifestly off-resonant for single-photon absorption
    processes, while multi-photon absorption processes are always
    off-resonant due to (i) the non-linearity (on the scale of kHz) of
    on-site electronic orbital excitation energies (resulting from
    non-uniform coupling constants $G_X$), combined with (ii) the
    ultranarrow linewidth (on the scale of mHz) of the interrogation
    laser.}

  We have also added, at the end of the first paragraph in our
  introduction that discusses our technical contributions, a sentence
  directing readers to the above summary of experimental procedures.
  Specifically, the text (page 4, line 31)

  \red{... through the exquisite capabilities with OLCs.}

  now reads

  \green{... through the exquisite capabilities with OLCs.  We briefly
    summarize the experimental procedures relevant to our work in
    Appendix A.}

  We have likewise directed readers to the new appendix at the
  beginning of the experiment/theory comparisons in section IV (page
  18), such that the first sentence of this section, previously

  \red{Current experiments with ultracold ${}^{87}$Sr on a lattice can
    coherently address ground states and their low-lying orbital
    excitations for up to five atoms per lattice site [34].}

  now reads

  \green{Current experiments with ultracold ${}^{87}$Sr on a lattice
    can coherently address ground states and their low-lying orbital
    excitations for up to five atoms per lattice site [34] (see
    Appendix A for a summary of relevant experimental procedures).}

  We hope that these additions provide a satisfactory description of
  the experimental procedures in ref.~[34], deeming our manuscript
  sufficiently self-contained for independent publication.

\item Concerning:

  \blue{(ii) The paragraph between Eq.(8) and Eq.(9) is unreadable and
    extremely hard to understand. E.g., the following sentences ``Even
    at zero temperature, therefore, the presence of excited motional
    states modifies many-body atomic ...'', or ``As our work is
    motivated by progress in experiments which initialize all atoms in
    identical (i.e. ground) orbital and motional states, to simplify
    our theory we assume that ...''.}

  \label{pt:confusion}

  We thank the referee for their careful reading of our manuscript,
  and for pointing out that our discussion in the paragraph between
  Eqs.~(8) and (9) is poorly worded.  The paragraph in question
  appears on page 7 lines 20--47:

  \red{We are interested in the ground and low-lying orbital
    eigenstates of several interacting atoms on a single lattice site,
    as such states can be readily prepared and coherently addressed in
    current experiments with ultracold atoms [34].  Although these
    experiments operate at temperatures sufficiently low for
    negligible thermal occupation of excited motional states, simply
    neglecting excited single-particle motional states fails to
    reproduce the observed spectra of multiply-occupied lattice sites.
    The reason for this failure lies in the fact that the bare
    two-body Hamiltonian $H_{\t{int}}$ in (7) is not diagonal in the
    basis of single-particle motional states.  Even at zero
    temperature, therefore, the presence of excited motional states
    modifies many-body atomic energy spectra when the magnitude of
    off-diagonal elements of $H_{\t{int}}$ approach the scale of
    low-lying single-particle motional state energy gaps.  To address
    this problem, we develop a low-energy effective theory for
    ultracold two-level fermionic atoms on a lattice.  As our work is
    motivated by progress in experiments which initialize all atoms in
    identical (i.e.~ground) orbital and motional states, to simplify
    our theory we assume that any $N$ atoms on a single lattice site
    occupy $N$ nuclear spin states.  Multiple occupation of a single
    nuclear spin state on the same lattice site is initially forbidden
    by fermionic statistics, and subsequently cannot be violated in
    the absence of tunneling and hyperfine coupling.}

  We have re-written this paragraph in its entirety, replacing it with
  the following text:

  \green{Current experiments with ${}^{87}$Sr can prepare up to five
    atoms atoms in the same (ground) orbital state on a single lattice
    site, and coherently address states with a single orbital
    excitation per lattice site (see Appendix A) [34].  The
    temperature of the atoms in these experiments is much smaller than
    single-atom motional excitation energies $\Delta_n\equiv E_n-E_0$
    for $n>0$, which implies a negligible thermal occupation of
    excited motional states.  A naive calculation of the multi-body
    orbital excitation spectrum, however, assuming that atoms only
    occupy motional ground states with spatial wavefunctions
    $\phi_{i,0}$, fails to reproduce the experimentally observed
    orbital excitation spectrum of multiply-occupied lattice sites.
    The reason that such a calculation fails to reproduce the observed
    excitation spectrum has to do with the fact that the two-body
    Hamiltonian $H_{\t{int}}$ in (7) is not diagonal in the basis of
    single-particle motional eigenstates (i.e.~$\phi_{in}$).  For
    interacting atoms, the true motional ground states are therefore
    admixtures of the non-interacting motional eigenstates; when the
    off-diagonal elements of $H_{\t{int}}$ are comparable in magnitude
    to the excitation energies $\Delta_n$, this admixture can have
    non-negligible weight outside the non-interacting motional
    ground-state subspace.  In order to compute the correct many-body
    orbital excitation spectrum, we develop a low-energy effective
    theory for such atoms.  As our work is motivated by experiments
    that initialize all atoms in identical (i.e.~ground) orbital and
    motional states, to simplify our theory we assume that any $N$
    atoms on a single lattice site occupy $N$ nuclear spin states.
    Ground-state preparation implies that multiple occupation of a
    single nuclear spin state on any given lattice site is initially
    forbidden by fermionic statistics, and subsequently cannot occur
    in the absence of inter-site effects or hyperfine coupling.}

  We hope that the new text is more readable and easier to understand.

\item Concerning:

  \blue{(iii) Eq. (10) and Eq. (42) are identical.}

  Indeed, Eq.~(10) (page 8) and Eq.~(42) (page 18) of our manuscript
  are identical.  The purpose of Eq.~(10) is to provide a preview and
  some brief discussion of one of our main results.  This same
  equation, however, is central to the discussions in section IV, so
  we provide this equation again in Eq.~(42) for quick reference.  We
  believe that the small cost in space is worth the added benefit of
  decoupling section IV (page 18) from section II (page 5), both (i)
  from a stylistic standpoint, and (ii) to save readers the need to
  locate Eq.~(10), and repeatedly flip back and forth between
  disparate sections of the manuscript.

  With these considerations in mind, we leave the choice of whether it
  is okay to have two identical equations up to the editors.  If the
  editors decide we must remove Eq.~(42), we would replace the
  following text (page 18, line 48):

  \red{Due to the SU($N$) symmetry of inter-atomic interactions,
    manifest in the fact that all coupling constants are independent
    of nuclear spin, we can write all effective $M$-body Hamiltonians
    which address these states in the form
    \begin{align}
      H_M = \sum_{\abs{\set{\mu_j}}=M}
      \p{U_{M,\g} \prod_{\alpha=1}^M \n_{\mu_\alpha,\g}
        + U_{M,+} \n_{\mu_1,\e} \prod_{\alpha=2}^M \n_{\mu_\alpha,\g}
        + U_{M,-} \c_{\mu_1,\g}^\dag \c_{\mu_2,\e}^\dag
        \c_{\mu_2,\g} \c_{\mu_1,\e} \prod_{\alpha=3}^M \n_{\mu_\alpha,\g}},
    \end{align}
    where $\n_{\mu s}\equiv \c_{\mu s}^\dag\c_{\mu s}$ is a number
    operator, and the coefficients $U_X$ can be determined from the
    coupling constants $G_Y$ and prefactors $\alpha_Z^{(p)}$ of the
    effective $M$-body Hamiltonians derived in section III (see
    Appendix G).}

  with

  \green{Due to the SU($N$) symmetry of inter-atomic interactions,
    manifest in the fact that all coupling constants are independent
    of nuclear spin, we can write all effective $M$-body Hamiltonians
    which address these states in the form of (10).}

  Note that we have not made this replacement in the re-submitted
  version of our manuscript, but will do so at the editors' request.

\item Concerning the interaction pseudopotential:

  \blue{(iv) It is well established that Eq.(2) is incorrect as the
    s-wave interaction in three-dimensions is described by a
    pseudo-potential and not a delta-function interaction. It is
    rather misleading that the authors to not point out the problem at
    the relevant position, but only afterwards require a
    renormalization of the divergencies without explaining the reason
    for their appearance.}

  There are two related matters to address in this comment:~one
  concerning the use of an interaction pseudo-potential, and another
  concerning divergences that appear in our effective theory.  We
  first address the question of a pseudo-potential.

  The referee is right that Eq.~(2) (page 6) is not the ``correct''
  form of the two-body interaction between two ultracold atoms.
  Rather, Eq.~(2) is a standard approximate form for two-body
  interactions (see, for example, Eqs.~(1) of
  refs.~[\href{https://www.nature.com/articles/nphys3061}{17}],
  [\href{https://iopscience.iop.org/article/10.1088/1367-2630/11/10/103033/meta}{21}],
  [\href{https://iopscience.iop.org/article/10.1088/1367-2630/14/5/053037/meta}{35}],
  [\href{https://iopscience.iop.org/article/10.1088/1367-2630/11/9/093022/meta}{37}],
  [\href{https://www.nature.com/articles/nphys1535}{41}],
  [\href{https://journals.aps.org/pra/abstract/10.1103/PhysRevA.87.033601}{52}],
  and
  [\href{https://link.springer.com/article/10.1023\%2FA\%3A1018705520999}{53}];
  and Eqs.~(4) and (5) of
  ref.~[\href{https://journals.aps.org/pra/abstract/10.1103/PhysRevA.90.043631}{36}])
  that appears in an effective field theory (EFT) description of
  low-energy scattering.  In general, an EFT description of
  interactions would also include terms with derivatives of field
  operators, taking the form of e.g.~Eqs.~(2) and (4) of
  ref.~[\href{https://journals.aps.org/pra/abstract/10.1103/PhysRevA.90.043631}{36}];
  scattering is thus generally momentum-dependent, implying the need
  to use a momentum-dependent (commonly referred to as an
  energy-dependent) interaction pseudo-potential.  The precise form of
  the dominant low-energy correction to the zero-momentum contact
  interaction in EFT is well-known (see
  refs.~[\href{https://journals.aps.org/pra/abstract/10.1103/PhysRevA.65.043613}{55}]
  and
  [\href{https://journals.aps.org/pra/abstract/10.1103/PhysRevA.59.1998}{57}]
  for studies of this correction, as well as
  refs.~[\href{https://iopscience.iop.org/article/10.1088/1367-2630/14/5/053037/meta}{35}],
  [\href{https://journals.aps.org/pra/abstract/10.1103/PhysRevA.90.043631}{36}],
  and
  [\href{https://iopscience.iop.org/article/10.1088/1367-2630/11/9/093022/meta}{37}]
  for their application to EFT calculations), and amounts to using an
  effective scattering length
  \begin{align*}
    a_{\t{eff}} = a - \f12 r_{\t{eff}} a^2 k^2,
  \end{align*}
  where $a$ is the zero-momentum scattering length, $r_{\t{eff}}$
  ($\sim a$ our case) is an effective range of the leading-order
  momentum-dependent interaction, and $k$ is the relative momentum of
  two atoms.  We include these corrections in our all of our numerical
  results, as discussed in the last paragraph of section III B (page
  13, line 24):

  \red{Second, the overview we provided in section II ...  neglects
    the effect of momentum-dependent two-body scattering. ... these
    interactions are third order in the coupling constants $G_X$.
    ... [and therefore] will not appear in any of our three- and
    four-body calculations ... The primary interest of our work
    ... concerns effective $M$-body interactions for $M\ge3$; we
    therefore defer the calculation of momentum-dependent two-body
    interactions to Appendix D.}

  A more detailed discussion of momentum-dependent corrections (and
  thereby our use of an energy-dependent pseudo-potential) is provided
  in Appendix D (page 36).

  Given the standard use of a zero-momentum contact interaction in the
  relevant literature, we did not consider it necessary to complicate
  our discussion in the overview of our manuscript (section II) by
  introducing an energy-dependent pseudo-potential.  In light of the
  referee's comment, however, we now see that our omission of this
  subtlety following Eq.~(2) may confuse some readers, or mislead them
  about the nature of two-body interactions.  In order to avoid such
  confusion without introducing major complications to our overview,
  we have therefore appended the following sentence to the paragraph
  containing Eq.~(5) (page 6, before line 36):

  \green{Note that (2) and (5) neglect momentum-dependent corrections
    to $s$-wave interactions, which are generally captured (in a
    low-energy limit) by the use of an energy-dependent interaction
    pseudo-potential [42].  These corrections will appear only at
    third order in our effective theory, as corrections to two-body
    interaction energies, and do not affect our discussions of
    multi-body interactions.  We therefore defer further discussion of
    momentum-dependent corrections to the end of section III B, and
    their calculation to Appendix E.}

  For an immediate reference on the subject matter, we have also added
  ref.~[55] (now ref.~[42] in the revised manuscript) to the end of
  the sentence on page 13, line 24, previously:

  \red{Second, the overview we provided in section II, and
    consequently our result in (17), neglects the effect of
    momentum-dependent two-body scattering.}

  and now:

  \green{Second, the overview we provided in section II, and
    consequently our result in (17), neglects the effect of
    momentum-dependent two-body scattering [42].}

  We hope that these additions clarify the point about our use of an
  energy-dependent pseudo-potential.

\item Concerning the second half of the previous comment by the
  referee, on divergences and renormalization:

  \blue{It is rather misleading that the authors to not point out the
    problem at the relevant position, but only afterwards require a
    renormalization of the divergencies without explaining the reason
    for their appearance.}

  We thank the referee for pointing out a potential point of confusion
  in our manuscript due to the appearance of divergent terms that must
  be regulated (in our case, by the introduction of counter-terms).
  To address this point of confusion, we have replaced the following
  text (page 12, line 25):

  \red{In general, the sums over excited states in the loop diagrams
    of (15) will diverge [35], which implies the need to renormalize
    the coupling constants in our theory.  Even if these diagrams do
    not diverge, we would like to express the effective two-body
    interactions in terms of net, physically observable two-body
    coupling constants as on the right-hand side of (15), rather than
    in terms of long sums at all orders of the bare coupling
    constants.  We therefore renormalize our coupling constants by
    introducing counter-terms $\tilde G_X$ into the effective
    interaction Hamiltonians via $G_X\to G_X+\tilde G_X$.  The values
    of these counter-terms are fixed by enforcing that $G_X$ is equal
    to the net two-body interaction strength.}

  with:

  \green{On physical grounds, the net effective two-body interaction
    must clearly be finite, but individual sums over the excited
    states in loop diagrams of (15) may generally diverge [35].  These
    divergences ultimately appear due to our use of effective field
    theory to describe inter-atomic interactions in (2), rather than a
    detailed microscopic description of two-atom scattering.
    Divergences of this sort are a generic feature of field theories,
    and can be dealt with using standard techniques, in particular
    through renormalization.  We therefore renormalize our coupling
    constants by introducing counter-terms $\tilde G_X$ into the
    effective interaction Hamiltonians via $G_X\to G_X+\tilde G_X$.
    The values of counter-terms can be fixed by enforcing that $G_X$
    is equal to the net two-body interaction strength, which has the
    added benefit of allowing us to express the effective two-body
    interactions in terms of net, physically observable two-body
    coupling constants on the right-hand side of (15), rather than in
    terms of long sums at all orders of the bare coupling constants.}

  We hope that this discussion helps clarify confusions regarding the
  appearance of divergences in our effective theory.

\item Concerning the effective theory corrections to many-body energy
  eigenstates:

  \blue{(v) On page 10: For the sentence ``Finally, the effective
    theory involves no corrections to the many-body energy
    eigenstates.'' the referee can again only guess, what the authors
    would like to express, but as it is written it is just plain
    wrong. The states between the non-interacting and the interacting
    theory are connected by a unitary transformation and are therefore
    not identical to the non-interacting ones. }

  We thank the referee for pointing out a subtle point that we need to
  clarify about effective theories constructed with a Schrieffer-Wolff
  transformation.  It is true, as the referee states, that the energy
  eigenstates $\set{\ket\alpha}$ of an interacting theory are not
  identical to those of a non-interacting theory,
  $\set{\ket{\alpha_0}}$, and that these sets of energy eigenstates
  are related by a unitary transformation
  $U=\sum_\alpha \op{\alpha_0}{\alpha}$.  While the non-interacting
  eigenstates $\set{\ket{\alpha_0}}$ are typically simple to identify,
  however, the interacting eigenstates $\set{\ket\alpha}$ are
  generally difficult to represent in the same basis as the
  Hamiltonian $H=H_0+H_{\t{int}}$ of the interacting theory, where in
  our case $H_0$ ($H_{\t{int}}$) contains only single-body (two-body)
  terms.  The purpose of an effective theory is then to construct an
  {\it effective Hamiltonian} $H_{\t{eff}}$ with two key
  properties:~(i) $H_{\t{eff}}$ is diagonal in the ``simple'' basis
  $\set{\ket{\alpha_0}}$, and (ii) the spectrum of $H_{\t{eff}}$ is
  identical to that of the true Hamiltonian $H$.  That is, if the
  non-interacting and interacting Hamiltonians respectively have
  decompositions $H_0 = \sum_\alpha E_\alpha \op{\alpha_0}$ and
  $H = \sum_\alpha \E_\alpha \op\alpha$, an effective Hamiltonian
  takes the form
  $H_{\t{eff}} = U H U^\dag = \sum_\alpha \E_\alpha \op{\alpha_0}$.
  The energy eigenstates of the effective theory are therefore
  identical to those of the non-interacting theory, namely
  $\set{\ket{\alpha_0}}$.

\end{enumerate}

\end{document}

### 2019-02-07 multi_body_interactions/papers/theory/reviews_NJP/referee_1_response.tex
updates/changes from responses to NJP referee 1

@@
 
-\item Concerning the comment:
+\item Concerning the generality of our multi-body theory:
 
@@
     restrict the analysis to a low density of excitations in the ``e''
-    orbital state. This restriction is never explained nor motivated.}
+    orbital state. This restriction is never explained nor motivated.
+    In addition, the restriction to a single occupation of the nuclear
+    spin state is never motivated or justified but just introduced at
+    some point. The referee understands, that these restrictions are
+    justified for the comparison with the experiment in
+    Ref.[34]. However, it should be clearly pointed, that the analysis
+    is not general and adapted to the description to these experiments
+    in the abstract and introduction. Furthermore, a short description
+    of the experiment would be mandatory for an independent
+    manuscript.}
 
-  \label{pt:general}
+  \label{pt:limitations}
 
-  The referee is correct that we do not ``derive a general low energy
-  theory for orbital and nuclear spin states''.  In our abstract, for
-  example, we merely claim that (page 1, line 27)
+  We thank the referee for pointing out the potential to mislead
+  readers on the generality of our theory.  The referee is correct
+  that we do not ``derive a general low energy theory for orbital and
+  nuclear spin states'', and we agree on the importance of making the
+  scope of our technical contributions clear.  In particular, our
+  theory does not consider multiple occupation of a nuclear spin state
+  on any lattice site.  To facilitate comparison with experiment, we
+  additionally consider a restriction of our theory to the space of
+  states with at most one orbital excitation per lattice site, as
+  e.g.~in Eq.~(10) (page 8) and section IV (page 18).  The full theory
+  developed in section III (page 8), however, is completely general
+  with respect to the number of orbital excitations per lattice site.
+  We also thank the referee for pointing out that a short description
+  of relevant experimental procedures would make our manuscript more
+  self-contained, and we recognize the value in doing so.
 
-  \red{Motivated in particular by recent experiments ... we develop a
-    low-energy effective theory ... which exhibits emergent multi-body
-    SU($N$)-symmetric interactions ...}
+  To clarify the limitations of our theory, we have added a sentence
+  to our abstract, such that the original text (page 1, line 34):
 
-  In our introduction, meanwhile, the very first sentence discussing
-  the technical contributions of our manuscript (page 4, line 19)
-  explicitly states that
+  \red{... nuclear spin levels.  We fully characterize...}
 
-  \red{In this work, we investigate the first experimental
-    capabilities with ultracold fermionic AEAs to prepare high-density
-    samples in a 3-D lattice with multiple occupation of individual
-    lattice sites [34].}
+  now reads:
 
-  We similarly begin both our abstract and introduction with a
-  discussion of specific experiments, and make frequent reference to
-  experimental developments and capabilities.  We thus believe that
-  our presentation is faithful to the technical content of our
-  manuscript; we do not make any claim of having developed a
-  completely general theory two-orbital multi-body SU($N$)-symmetric
-  interactions.  We claim only to have developed a low-energy theory
-  that (i) is relevant for current experiments, and (ii) exhibits
-  multi-body SU($N$)-symmetric interactions.  For these reasons, we do
-  indeed introduce several approximations even in the first paragraph
-  of our theory overview (section II, page 5).  We believe that
-  theoretical work motivated by specific experiments is generally
-  justified in making approximations that are relevant to the
-  parameter regimes of these experiments.
+  \green{... nuclear spin levels.  Our theory is limited to the
+    experimental regime of (i) a deep lattice, with (ii) at most one
+    atom occupying each nuclear spin state on any lattice site, where
+    the latter restriction is a consequence of initial ground-state
+    preparation.  We fully characterize...}
 
-  Nonetheless, we fully recognize and agree with the referee on the
-  importance of making the scope of our technical contributions clear
-  (as the referee comments later in point \ref{pt:restrictions}).  The
-  fact that the referee at any point believed we will present a
-  general theory without applying several experimentally-motivated
-  approximations clearly signifies that we need to be more explicit in
-  our aims.  Consequently, we have modified the second sentence of our
-  original abstract, from (page 1, line 27)
-
-  \red{Motivated in particular by recent experiments ...}
-
-  to
-
-  \green{In order to analyze recent experiments ...}
-
-  We have also clarified an important point in our introduction,
-  replacing the following text (page 4, line 22)
+  In the introduction, we have modified the following sentence (in the
+  first paragraph discussing our technical contributions) (page 4,
+  line 22):
 
@@
 
-  with
+  which now reads:
 
@@
 
-  The point on ground-state preparation in the deep lattice limit is
-  important for a restriction we make later concerning single
-  occupation of nuclear spin states on individual lattice sites.  This
-  restriction (discussed in point \ref{pt:single_nuclear} below)
-  together with the experimentally-motivated approximations made in
-  the first paragraph of the theory overview (section II, page 5)
-  constitute all key restrictions of our theory.
+  and we have appended an additional sentence to the same paragraph,
+  previously ending with the text (page 4, line 31):
 
-  We hope that these changes, together with those discussed in the
-  examples below (points \ref{pt:orbital_state} and
-  \ref{pt:single_nuclear}), clarify the motivation and scope of our
-  manuscript, such that a reader is not led to believe that we present
-  a general theory without applying several well-motivated
-  approximations that are relevant to current experiments.
+  \red{... through the exquisite capabilities with OLCs.}
 
-\item Concerning the specific example:
+  and now:
 
-  \blue{However, in Eq.(10) the authors suddenly restrict the analysis
-    to a low density of excitations in the ``e'' orbital state. This
-    restriction is never explained nor motivated.}
+  \green{... through the exquisite capabilities with OLCs.  Our theory
+    is limited to the experimental regime of at most one atom
+    occupying each nuclear spin state on any lattice site, which is a
+    consequence initial ground-state preparation.}
 
-  \label{pt:orbital_state}
+  We have also rearranged and rewritten some of the following
+  paragraph (page 4, line 33):
 
-  It is true, as the referee points out, that in Eq.~(10) (page 8) we
-  present an effective Hamiltonian addressing only states with at most
-  one orbital excitation (``e'') per lattice site.  We motivate this
-  restriction immediately after Eq.~(8) in our theory overview, as
-  soon as we begin a technical discussion of multi-body interactions
-  (page 7, line 20):
+  \red{Though effective multi-body interactions have previously been
+    studied in the context of harmonically [35, 36] and
+    lattice-confined [37] neutral bosons prepared in a single
+    hyperfine state, our work deals for the first time with fermions
+    that have internal degrees of freedom and multiple collisional
+    parameters.  Nonetheless, due to the SU($N$) symmetry of these
+    collisions, we are able to find a simple way to express effective
+    multi-body Hamiltonians and fully characterize their low-lying
+    many-body energy eigenstates.  While some past work has detected
+    experimental signatures of multi-body interactions in the form of
+    quantum phase revivals [38], we perform a direct comparison of the
+    many-body interaction energies predicted by our low-energy
+    effective theory with experimental measurements of
+    density-dependent atomic energy spectra performed in ref.~[34],
+    similarly to the measurements with bosons performed in ref.~[39].}
 
-  \red{We are interested in the ground and low-lying orbital
-    eigenstates of several interacting atoms on a single lattice site,
-    as such states can be readily prepared and coherently addressed in
-    current experiments with ultracold atoms [34].}
+  which now reads:
 
-  We acknowledge that the referee found the paragraph containing this
-  sentence confusing, and have consequently rewritten the paragraph
-  between Eqs.~(8) and (9) entirely (see point \ref{pt:confusion}).
-  In any case, our actual results in section III (page 8) take the
-  general form of Eq.~(9) (page 7), and make no restrictions on the
-  number of orbital excitations per lattice site.  The expression in
-  Eq.~(10) is merely a simplified restriction of our technical
-  results, made (i) to highlight important features of the low-lying
-  orbital excitations in our multi-body SU($N$) theory, and (ii) to
-  facilitate comparison with optical lattice clock experiments.
+  \green{Though effective multi-body interactions have previously been
+    studied in the context of harmonically [35, 36] and
+    lattice-confined [37] neutral bosons prepared in a single
+    hyperfine state, our work deals for the first time with fermions
+    that have internal degrees of freedom and multiple collisional
+    parameters. We perform a direct comparison between the many-body
+    interaction energies predicted by our low-energy effective theory
+    and the experimental measurements of the density-dependent orbital
+    excitation spectra performed in ref.~[34], similarly to the
+    measurements with bosons performed in ref.~[38]. To facilitate
+    this comparison of excitation spectra and characterize the
+    low-lying excitations in our effective theory, we consider a
+    restriction of the theory to states with at most one orbital
+    excitation per lattice site, and find that the SU($N$) symmetry of
+    atomic collisions allow effective multi-body interactions to take
+    a remarkably simple form. We note that past work has also detected
+    experimental signatures of multi-body interactions in the form of
+    quantum phase revivals [39].}
 
-  To clarify this point, we have made more explicit the fact that our
-  actual results take the form of Eq.~(9), and added additional
-  discussion of Eq.~(10).  Specifically, we have replaced the
-  following text (page 7, line 48):
+  We also clarify the limitations of our theory in the overview
+  between Eqs.~(8) and (9), as discussed later in point
+  \ref{pt:confusion}, and have replaced the following text (page 7,
+  line 48):
 
@@
     where $H_M^{(p)}$ is an $M$-body Hamiltonian of order $p$ in the
-    coupling constants $G^{qr}_{st}$, $I$ is the total nuclear spin of
-    each atom (e.g.~$I=9/2$ for ${}^{87}$Sr), and the sum terminates
-    at $2I+1$ because this is the largest number of atoms which may
+    coupling constants $G_X$, $I$ is the total nuclear spin of each
+    atom (e.g.~$I=9/2$ for ${}^{87}$Sr), and the sum terminates at
+    $2I+1$ because this is the largest number of atoms which may
     initially occupy a single lattice site.  We explicitly compute all
@@
     $p=3$, yielding effective two-, three-, and four-body
-    interactions.  To facilitate understanding and comparison with
-    experimental measurements of density-dependent atomic energy
-    spectra [34], in which states with multiple orbital excitations on
-    a single lattice site are spectroscopically off-resonant, we also
-    consider a restriction of our theory to states with at most one
-    orbital excitation per lattice site.  In this case, we find that
-    SU($N$) symmetry allows us to express multi-body Hamiltonians in
-    the simple form}
+    interactions.  To (i) facilitate a comparison with the
+    experimental measurements of many-body orbital excitation spectra
+    performed in ref.~[34] and (ii) characterize the low-lying
+    excitations in our effective theory, we also consider a
+    restriction of the multi-body Hamiltonians $H_M$ to states with at
+    most one orbital excitation per lattice site.  We find that the
+    SU($N$) symmetry of atomic collisions allows us to express
+    multi-body Hamiltonians in the simple form}
 
-  We have also modified the following sentence in our summary (page
-  28, line 43):
+  In section IV (page 18), we have changed the first sentence from
+  (page 18, line line 45):
 
-  \red{The SU($N$) symmetry of $M$-body Hamiltonians allowed us to
+  \red{Current experiments with ultracold ${}^{87}$Sr on a lattice can
+    coherently address ground states and their low-lying orbital
+    excitations for up to five atoms per lattice site [34].  Due to
+    the SU($N$) symmetry of inter-atomic interactions, manifest in the
+    fact that all coupling constants are independent of nuclear spin,
+    we can write all effective $M$-body Hamiltonians which address
+    these states in the form}
+
+  to:
+
+  \green{Current experiments with ultracold ${}^{87}$Sr on a lattice
+    can coherently address ground states and single orbital
+    excitations of up to five atoms per lattice site [34].  Due to the
+    SU($N$) symmetry of inter-atomic interactions, manifest in the
+    fact that all coupling constants are independent of nuclear spin,
+    a restriction of the $M$-body Hamiltonians $H_M=\sum_p H_M^{(p)}$
+    to the subspace of experimentally addressed states takes the form}
+
+  Finally, we have modified the following text in our summary (page
+  28, line 40):
+
+  \red{Working in the deep-lattice limit, we have derived a low-energy
+    effective theory of such atoms. This theory exhibits emergent
+    multi-body interactions that inherit the SU($N$) symmetry of the
+    atoms' bare, two-body interactions. The SU($N$) symmetry of
+    $M$-body Hamiltonians in the effective theory has allowed us to
     express them in a simple form, and fully characterize their
@@
 
-  \green{When considering the low-lying orbitalexcitations of our
-    effective theory, the SU($N$) symmetry of $M$-body Hamiltonians
-    allowed us to express them in a simple form, and to fully
-    characterize their eigenstates and spectra.}
-
-  Finally, the new appendix we have added as a part of point
-  \ref{pt:description} (below) contains additional discussion of the
-  restriction to at most one orbital excitation per lattice site.
+  \green{Working in the deep-lattice limit and the experimental regime
+    of at most one atom occupying each nuclear spin state on any
+    lattice site, we have derived a low-energy effective theory for
+    such atoms.  Our theory exhibits emergent multi-body interactions
+    that inherit the SU($N$) symmetry of atoms' bare, two-body
+    interactions.  Considering a restriction of our theory to the
+    subspace of at most one orbital excitation per lattice site, we
+    found that the SU($N$) symmetry of $M$-body Hamiltonians allowed
+    us to express them in a simple form, and to fully characterize
+    their eigenstates and spectra.}
 
   We hope that these changes clarify the scope of our work, in
-  particular highlighting that the theory developed in section III
-  applies to any number of orbital excitations, despite the
-  restriction to the single-excitation case for the sake of comparison
-  with experiments in section IV (page 18).
+  particular highlighting that (i) our theory is limited to the case
+  of at most one atom per nuclear spin state per lattice site, and
+  (ii) the theory in section III applies to any number of orbital
+  excitations, despite the restriction to the single-excitation case
+  for the sake of comparison with experiments in section IV.
 
-\item Concerning the next example:
+  Concerning a description of the experiment in ref.~[34], we have
+  added a summary of the relevant experimental procedures to the
+  beginning of section II, such that the previous opening (page 5,
+  line 19):
 
-  \blue{In addition, the restriction to a single occupation of the
-    nuclear spin state is never motivated or justified but just
-    introduced at some point.}
+  \red{Consider a collection of fermionic atoms with total nuclear
+    spin $I$ and two atomic orbital states loaded into a
+    translationally invariant 3-D optical lattice.  While an external
+    trapping potential will generally break discrete translational
+    symmetry of the lattice, any background inhomogeneity can be made
+    negligible by spectroscopically addressing a sufficiently small
+    region of the lattice [34].  Throughout this paper, we work
+    strictly in the deep-lattice regime with negligible tunneling
+    between lattice sites.  We also assume a so-called
+    ``magic-wavelength'' lattice for which both atomic orbital states
+    experience identical lattice potentials [40].}
 
-  \label{pt:single_nuclear}
-
-  The referee is correct that we restrict our theory to the case of
-  single occupation of nuclear spin states on any given lattice site.
-  We do so because (i) this regime is precisely that relevant to
-  current experiments with ultracold atoms, which begin with
-  ground-state preparation of the atoms, and (ii) this restriction
-  greatly simplifies our theory.  We explained both the motivation and
-  the justification for this restriction in the overview of our theory
-  (page 7, line 40):
-
-  \red{As our work is motivated by progress in experiments which
-    initialize all atoms in identical (i.e.~ground) orbital and
-    motional states, to simplify our theory we assume that any $N$
-    atoms on a single lattice site occupy $N$ nuclear spin states.
-    Multiple occupation of a single nuclear spin state on the same
-    lattice site is initially forbidden by fermionic statistics, and
-    subsequently cannot be violated in the absence of tunneling and
-    hyperfine coupling.}
-
-  Deferring the referee's later comment that this text was confusing
-  to point \ref{pt:confusion}, we emphasize the fact that our theory
-  is the first to include hyperfine structure when considering
-  multi-body interactions of ultracold atoms on a lattice.  Of course,
-  a more general theory relaxing the above restriction (i.e.~on the
-  occupation of nuclear spin states) will eventually be necessary for
-  future experiments with a shallower lattice or additional control
-  fields.  Such a general theory will need to consider inter-site
-  effects or hyperfine coupling (or, less likely, non-ground-state
-  preparation protocols), as these are the mechanisms by which a
-  nuclear spin state can become multiply-occupied on a single lattice
-  site.  Given the that these mechanisms (i) are largely irrelevant
-  for the experiments that prompted this work, and (ii) introduce
-  major complications into the study of SU($N$) multi-body
-  interactions, we believe that it is appropriate to leave such
-  considerations for future work.
-
-  For clarification, we have modified the sentence on page 7, line 43
-  from
-
-  \red{Multiple occupation of a single nuclear spin state on the same
-    lattice site is initially forbidden by fermionic statistics, and
-    subsequently cannot be violated in the absence of tunneling and
-    hyperfine coupling.}
-
-  to
-
-  \green{Ground-state preparation implies that multiple occupation of
-    a single nuclear spin state on any given lattice site is initially
-    forbidden by fermionic statistics, and subsequently cannot occur
-    in the absence of inter-site effects or hyperfine coupling.}
-
-  We hope that this modification, together with the modifications
-  mentioned in point \ref{pt:general}, clarifies our justification for
-  restricting to the case of single occupation of nuclear spin states.
-
-\item Concerning the following comments:
-
-  \blue{The referee understands, that these restrictions are justified
-    for the comparison with the experiment in Ref.[34]. However, it
-    should be clearly pointed, that the analysis is not general and
-    adapted to the description to these experiments in the abstract
-    and introduction.}
-
-  \label{pt:restrictions}
-
-  We thank the referee for pointing out the potential to mislead
-  readers on the generality of our theory, and hope that the new
-  version of our manuscript does a better job of calibrating the
-  expectations of a reader with regard to restrictions and
-  approximations that we make.  In particular, we hope that the
-  current text makes clear to readers that we are developing a theory
-  for the purpose of understanding current optical lattice clock
-  experiments, and not to understand SU($N$) multi-body interactions
-  in full generality.
-
-\item Concerning the description of the experiment with which we
-  compare:
-
-  \blue{Furthermore, a short description of the experiment would be
-    mandatory for an independent manuscript.}
-
-  \label{pt:description}
-
-  We thank the referee for pointing out that a description of relevant
-  experimental procedures would make our manuscript more
-  self-contained, and we recognize the value in doing so.  We have
-  therefore added a new appendix to our manuscript, summarizing all
-  relevant experimental procedures:
-
-  \green{{\bf Appendix A:~Summary of experimental procedures}}
+  now reads:
 
   \green{The work in this paper is closely tied to the experimental
-    work in ref.~[34]; for quick reference, we here provide a short
-    summary of the experimental procedures used therein.  The
-    experiment begins by preparing a degenerate gas of $10^4$-$10^5$
-    (fermionic) ${}^{87}$Sr atoms, in a uniform mixture of their ten
-    nuclear spin states, at $\sim0.1$ of their fermi temperature (10
-    to 20 nanokelvin) [8, 52].  This gas is loaded into a primitive
-    cubic optical lattice at the ``magic wavelength'' for which both
-    ground (${}^1S_0$) and first-excited (${}^3P_0$) electronic
-    orbital states of the atoms experience the same lattice potential.
-    Lattice depths along the principal axes of the lattice are
-    comparable in magnitude, with a geometric mean that can be varied
-    from 30 to 80 $E_{\t{R}}$, where
+    work in ref.~[34]; we begin with a short summary of the relevant
+    experimental procedures therein.  The experiment begins by
+    preparing a degenerate gas of $10^4$-$10^5$ (fermionic)
+    ${}^{87}$Sr atoms in a uniform mixture of their ten nuclear spin
+    states and at $\sim0.1$ of their fermi temperature ($\sim10$
+    nanokelvin) [8, 40].  This gas is loaded into a primitive cubic
+    optical lattice at the ``magic wavelength'' for which both ground
+    (${}^1S_0$) and first-excited (${}^3P_0$) electronic orbital
+    states of the atoms experience the same lattice potential [41].
+    Lattice depths along the principal axes of the lattice are roughly
+    equal in magnitude, with a geometric mean that can be varied from
+    30 to 80 $E_{\t{R}}$, where
     $E_{\t{R}}\approx3.5\times2\pi~\t{kHz}$ is the lattice photon
-    recoil energy of the atoms.  These lattice depths are sufficiently
-    large as to neglect tunneling on the time scales relevant to the
-    experiment.  Atoms are then addressed by an external (``clock'')
-    laser with an ultranarrow (26 mHz) linewidth, detuned by $\Delta$
-    from the single-atom ${}^1S_0-{}^3P_0$ transition frequency.  When
-    $\Delta=0$, non-interacting atoms in singly-occupied lattice sites
-    thus undergo Rabi oscillations.  After a chosen time, the
-    experiment removes all ground-state (${}^1S_0$) atoms from the
-    lattice and uses absorption imaging to identify the remaining
-    excited-state (${}^3P_0$) atoms.  As discussed in section IV, a
-    similar procedure with $\Delta\ne0$ leads to the spectrum shown in
-    figure 3, thereby constituting a measurement of the many-body
-    atomic excitation spectrum.  The many-body atomic excitation
-    spectrum is different from the single-body atomic excitation
-    spectrum due to the presence of inter-atomic interactions.  We
-    note that this spectroscopic protocol addresses only
-    singly-excited orbital states of any lattice site.  Doubly-excited
-    states are manifestly off-resonant for single-photon absorption
-    processes, while multi-photon absorption processes are always
-    off-resonant due to (i) the non-linearity (on the scale of kHz) of
-    on-site electronic orbital excitation energies (resulting from
-    non-uniform coupling constants $G_X$), combined with (ii) the
-    ultranarrow linewidth (on the scale of mHz) of the interrogation
-    laser.}
+    recoil energy of the atoms (with the reduced Planck constant
+    $\hbar=1$ throughout this paper).  These lattice depths are
+    sufficiently large as to neglect tunneling on the time scales
+    relevant to the experiment.  The temperature of the atoms is also
+    low enough to neglect thermal occupation of states outside the
+    ground-state manifold.}
 
-  We have also added, at the end of the first paragraph in our
-  introduction that discusses our technical contributions, a sentence
-  directing readers to the above summary of experimental procedures.
-  Specifically, the text (page 4, line 31)
+  \green{Once loaded into an optical lattice lattice, atoms are
+    addressed by an external (``clock'') interrogation laser with an
+    ultranarrow (26 mHz) linewidth, detuned by $\Delta$ from the
+    single-atom ${}^1S_0-{}^3P_0$ transition frequency $\omega_0$.
+    After a fixed interrogation time, the experiment turns off the
+    interrogation laser, removes all ground-state (${}^1S_0$) atoms
+    from the lattice, and uses absorption imaging to count any
+    remaining excited-state (${}^3P_0$) atoms.  Non-interacting atoms
+    in singly-occupied lattice sites undergo Rabi oscillations when
+    $\Delta=0$ during the interrogation time.  The orbital excitation
+    spectrum of multiply-occupied lattice sites, however, is shifted
+    by inter-atomic interactions, which results in spectroscopic peaks
+    (i.e.~local maxima in excited-state atom counts) away from
+    $\Delta=0$.  A sweep across different detunings $\Delta$ (on the
+    scale of inter-atomic interaction energies) thus constitutes a
+    measurement of the many-body orbital excitation spectrum.  We note
+    that this spectroscopic protocol addresses only singly-excited
+    orbital states of any lattice site.  Doubly-excited states are
+    manifestly off-resonant for single-photon absorption processes as
+    $\abs{\Delta}\ll\omega_0$, while multi-photon absorption processes
+    are always off-resonant due to (i) the non-linearity ($\sim$kHz)
+    of orbital excitation energies for a collection of interacting
+    atoms, and (ii) the ultranarrow linewidth ($\sim$mHz) of the
+    interrogation laser.}
 
-  \red{... through the exquisite capabilities with OLCs.}
+  \green{Consider now a collection of fermionic atoms with total
+    nuclear spin $I$ and two atomic orbital states loaded into a
+    translationally invariant 3-D optical lattice.  While an external
+    trapping potential will generally break discrete translational
+    symmetry of the lattice, any background inhomogeneity can be made
+    negligible by spectroscopically addressing a sufficiently small
+    region of the lattice [34].  Throughout this paper, we work
+    strictly in the deep-lattice regime with negligible tunneling
+    between lattice sites, and assume that both atomic orbital states
+    experience identical lattice potentials.}
 
-  now reads
+  To reflect the addition of new text, we have modified the
+  description of section II on page 4, line 52, previously:
 
-  \green{... through the exquisite capabilities with OLCs.  We briefly
-    summarize the experimental procedures relevant to our work in
-    Appendix A.}
+  \red{In section II we provide an overview of the relevant one- and
+    two-body physics of ultracold atoms on a lattice, and summarize
+    our main result of effective multi-body interactions between these
+    atom.}
 
-  We have likewise directed readers to the new appendix at the
-  beginning of the experiment/theory comparisons in section IV (page
-  18), such that the first sentence of this section, previously
+  and now:
 
-  \red{Current experiments with ultracold ${}^{87}$Sr on a lattice can
-    coherently address ground states and their low-lying orbital
-    excitations for up to five atoms per lattice site [34].}
+  \green{In section II we summarize the experimental procedures
+    relevant to our work, provide an overview one- and two-body
+    physics of ultracold atoms on a lattice, and preview our main
+    technical results.}
 
-  now reads
+  We have also removed the now redundant parenthetical text on page 5,
+  line 45:
 
-  \green{Current experiments with ultracold ${}^{87}$Sr on a lattice
-    can coherently address ground states and their low-lying orbital
-    excitations for up to five atoms per lattice site [34] (see
-    Appendix A for a summary of relevant experimental procedures).}
+  \red{(with the reduced Planck constant $\hbar=1$ throughout this
+    paper)}
 
-  We hope that these additions provide a satisfactory description of
-  the experimental procedures in ref.~[34], deeming our manuscript
-  sufficiently self-contained for independent publication.
+  We hope that the additional text in section II provides a
+  satisfactory description of the relevant experimental procedures in
+  ref.~[34], deeming our manuscript sufficiently self-contained for
+  independent publication.
 
@@
   Eqs.~(8) and (9) is poorly worded.  The paragraph in question
-  appears on page 7 lines 20--47:
+  appears on page 7, lines 20--47:
 
@@
     site, and coherently address states with a single orbital
-    excitation per lattice site (see Appendix A) [34].  The
-    temperature of the atoms in these experiments is much smaller than
-    single-atom motional excitation energies $\Delta_n\equiv E_n-E_0$
-    for $n>0$, which implies a negligible thermal occupation of
-    excited motional states.  A naive calculation of the multi-body
-    orbital excitation spectrum, however, assuming that atoms only
-    occupy motional ground states with spatial wavefunctions
-    $\phi_{i,0}$, fails to reproduce the experimentally observed
-    orbital excitation spectrum of multiply-occupied lattice sites.
-    The reason that such a calculation fails to reproduce the observed
-    excitation spectrum has to do with the fact that the two-body
-    Hamiltonian $H_{\t{int}}$ in (7) is not diagonal in the basis of
-    single-particle motional eigenstates (i.e.~$\phi_{in}$).  For
-    interacting atoms, the true motional ground states are therefore
-    admixtures of the non-interacting motional eigenstates; when the
-    off-diagonal elements of $H_{\t{int}}$ are comparable in magnitude
-    to the excitation energies $\Delta_n$, this admixture can have
-    non-negligible weight outside the non-interacting motional
-    ground-state subspace.  In order to compute the correct many-body
-    orbital excitation spectrum, we develop a low-energy effective
-    theory for such atoms.  As our work is motivated by experiments
-    that initialize all atoms in identical (i.e.~ground) orbital and
-    motional states, to simplify our theory we assume that any $N$
-    atoms on a single lattice site occupy $N$ nuclear spin states.
-    Ground-state preparation implies that multiple occupation of a
-    single nuclear spin state on any given lattice site is initially
-    forbidden by fermionic statistics, and subsequently cannot occur
-    in the absence of inter-site effects or hyperfine coupling.}
+    excitation per lattice site [34].  At ultracold temperatures well
+    below the single-atom motional excitation energies
+    $\Delta_n\equiv E_n-E_0$ for $n>0$, atoms only occupy their
+    motional ground state in the lattice.  For this reason, it is
+    common to map the description of these atoms onto a single-band
+    Hubbard model that captures all dynamics within the subspace of
+    motional ground states, i.e.~with single-atom wavefunctions
+    $\phi_{i,0}$.  When enough atoms occupy the same lattice site,
+    however, their interaction energy can approach the scale of the
+    motional excitation energies $\Delta_n$.  In this case, normally
+    far-off-resonant terms in the interaction Hamiltonian of (7) that
+    create atoms in excited motional states,
+    e.g.~$\sim\c_{n\mu s}^\dag \c_{0,\nu t}^\dag \c_{0,\nu r}
+    \c_{0,\mu q}$ with $n>0$, become relevant.  The true motional
+    ground state of a collection of interacting atoms is then an
+    admixture of the non-interacting motional eigenstates, and a naive
+    Hubbard model that assumes atomic wavefunctions $\phi_{i,0}$ will
+    fail to reproduce the interacting atoms' orbital excitation
+    spectrum.}
+
+  \green{In order to understand interacting atoms' orbital excitation
+    spectrum in the context of a zero-temperature Hubbard-type model
+    that eliminates non-dynamical (i.e.~ground-state) motional degrees
+    of freedom, we develop a low-energy effective theory of
+    interacting AEAs in a deep lattice.  To simplify our theory, we
+    assume that any $N$ atoms on a single lattice site occupy $N$
+    nuclear spin states.  Ground-state preparation of the atoms in an
+    experiment implies that multiple occupation of a single nuclear
+    spin state on any given lattice site is initially forbidden by
+    fermionic statistics, and subsequently cannot occur in the absence
+    of inter-site effects or hyperfine coupling between nuclear spin
+    states.}
+
+  Accordingly, we have modified language in section III that makes
+  reference to the re-written paragraph, changing (page 9, line 45):
+
+  \red{The naive idea of neglecting all excited states and using
+    $H_{\t{int}}$ directly to describe the spectrum of single-particle
+    motional ground states is thus equivalent to truncating our
+    expansion for $H_{\t{int}}^{\t{eff}}$ at first order.}
+
+  to:
+
+  \green{Writing down a single-band Hubbard model that simply neglects
+    excited atomic motional states and uses $H_{\t{int}}$ directly to
+    describe the orbital spectrum of interacting atoms is thus
+    equivalent to truncating our expansion for $H_{\t{int}}^{\t{eff}}$
+    at first order.}
 
@@
 
-  Indeed, Eq.~(10) (page 8) and Eq.~(42) (page 18) of our manuscript
-  are identical.  The purpose of Eq.~(10) is to provide a preview and
-  some brief discussion of one of our main results.  This same
-  equation, however, is central to the discussions in section IV, so
-  we provide this equation again in Eq.~(42) for quick reference.  We
-  believe that the small cost in space is worth the added benefit of
-  decoupling section IV (page 18) from section II (page 5), both (i)
-  from a stylistic standpoint, and (ii) to save readers the need to
-  locate Eq.~(10), and repeatedly flip back and forth between
-  disparate sections of the manuscript.
+  As the referee points out, Eqs.~(10) and (42) in the original
+  version of our manuscript are identical.  In order to avoid
+  redundancy while still providing a preview of our main results, we
+  have replaced the following text in section II (page 8, line 10):
 
-  With these considerations in mind, we leave the choice of whether it
-  is okay to have two identical equations up to the editors.  If the
-  editors decide we must remove Eq.~(42), we would replace the
-  following text (page 18, line 48):
+  \red{\begin{align} H_M = \sum_{\abs{\set{\mu_j}}=M} \p{U_{M,\g}
+        \prod_{\alpha=1}^M \n_{\mu_\alpha,\g} + U_{M,+} \n_{\mu_1,\e}
+        \prod_{\alpha=2}^M \n_{\mu_\alpha,\g} + U_{M,-}
+        \c_{\mu_1,\g}^\dag \c_{\mu_2,\e}^\dag \c_{\mu_2,\g}
+        \c_{\mu_1,\e} \prod_{\alpha=3}^M \n_{\mu_\alpha,\g}},
+      \tag{10}
+    \end{align}
+    where $\c_{\mu s}\equiv\c_{0,\mu s}$ is a fermionic annihilation
+    operator for an atom in the motional ground state with nuclear
+    spin $\mu$ and orbital state $s$;
+    $\n_{\mu s}\equiv \c_{\mu s}^\dag \c_{\mu s}$ is a number
+    operator; and the sum is performed over all choices of nuclear
+    spins $\mu_j$ with $j=1,2,\cdots,M$ for which all $\mu_j$ are
+    distinct, or equivalently all choices of $\mu_j$ for which the set
+    $\set{\mu_j}$ contains $M$ elements, for a total of
+    ${2I+1\choose M}\times\p{M!}$ nuclear spin combinations.  The
+    coefficients $U_X$ can be expressed in terms of the coupling
+    constants of the bare two-body interaction Hamiltonian (see
+    Appendix F).  The key feature of the $M$-body interactions in (10)
+    is that they take a form identical to the two-body interactions
+    between motional ground-state atoms, but with the addition of
+    $M-2$ spectator atoms.}
 
-  \red{Due to the SU($N$) symmetry of inter-atomic interactions,
-    manifest in the fact that all coupling constants are independent
-    of nuclear spin, we can write all effective $M$-body Hamiltonians
-    which address these states in the form
-    \begin{align}
+  by:
+
+  \green{\begin{align}
+      H_M = \sum_{\abs{\set{\mu_j}}=M}
+      H_2^{(\mu_1,\mu_2)} \prod_{\alpha=3}^M \n_{\mu_\alpha,\g},
+      \tag{10}
+    \end{align}
+    where $H_2^{(\mu_1,\mu_2)}$ is a two-body Hamiltonian addressing
+    atoms with nuclear spin $\mu_1,\mu_2$;
+    $\n_{\mu s}=\c_{\mu s}^\dag\c_{\mu s}$ is a number operator for
+    atoms with nuclear state $\mu$ and orbital state $s$; and the sum
+    is performed over all choices of nuclear spins $\mu_j$ with
+    $j=1,2,\cdots,M$ for which all $\mu_j$ are distinct, or
+    equivalently all choices of $\mu_j$ for which the set
+    $\set{\mu_j}$ contains $M$ elements, for a total of
+    ${2I+1\choose M}\times\p{M!}$ nuclear spin combinations.  The key
+    feature of the $M$-body interactions in (10) is that they
+    ultimately take the same form as two-body interactions, but with
+    the addition of $M-2$ spectator atoms.}
+
+  and have modified Eq.~(42) to take the explicit form advertised in
+  the new Eq.~(10), changing (page 18, line 54):
+
+  \red{\begin{align}
       H_M = \sum_{\abs{\set{\mu_j}}=M}
@@
         \c_{\mu_2,\g} \c_{\mu_1,\e} \prod_{\alpha=3}^M \n_{\mu_\alpha,\g}},
-    \end{align}
-    where $\n_{\mu s}\equiv \c_{\mu s}^\dag\c_{\mu s}$ is a number
-    operator, and the coefficients $U_X$ can be determined from the
-    coupling constants $G_Y$ and prefactors $\alpha_Z^{(p)}$ of the
-    effective $M$-body Hamiltonians derived in section III (see
-    Appendix G).}
+      \tag{42}
+    \end{align}}
 
-  with
+  to:
 
-  \green{Due to the SU($N$) symmetry of inter-atomic interactions,
-    manifest in the fact that all coupling constants are independent
-    of nuclear spin, we can write all effective $M$-body Hamiltonians
-    which address these states in the form of (10).}
+  \green{\begin{align}
+      H_M = \sum_{\abs{\set{\mu_j}}=M}
+      \p{U_{M,\g} \n_{\mu_1,\g} \n_{\mu_2,\g}
+        + U_{M,+} \n_{\mu_1,\e} \n_{\mu_2,\g}
+        + U_{M,-} \c_{\mu_1,\g}^\dag \c_{\mu_2,\e}^\dag
+        \c_{\mu_2,\g} \c_{\mu_1,\e}}
+      \prod_{\alpha=3}^M \n_{\mu_\alpha,\g},
+      \tag{42}
+    \end{align}}
 
-  Note that we have not made this replacement in the re-submitted
-  version of our manuscript, but will do so at the editors' request.
-
-\item Concerning the interaction pseudopotential:
+\item Concerning the use of an interaction pseudopotential:
 
@@
 
-  There are two related matters to address in this comment:~one
-  concerning the use of an interaction pseudo-potential, and another
-  concerning divergences that appear in our effective theory.  We
-  first address the question of a pseudo-potential.
+  We thank the referee for their attention to detail, and for pointing
+  out the potential to mislead readers with respect to the nature of
+  $s$-wave interactions by our expression of Eq.~(2).  We have
+  therefore added the following paragraph to section II of our
+  manuscript, following the re-statement of Eq.~(2) in Eq.~(5)
+  (i.e.~the following text has been inserted in between lines 35 and
+  36 of page 6):
 
-  The referee is right that Eq.~(2) (page 6) is not the ``correct''
-  form of the two-body interaction between two ultracold atoms.
-  Rather, Eq.~(2) is a standard approximate form for two-body
-  interactions (see, for example, Eqs.~(1) of
-  refs.~[\href{https://www.nature.com/articles/nphys3061}{17}],
-  [\href{https://iopscience.iop.org/article/10.1088/1367-2630/11/10/103033/meta}{21}],
-  [\href{https://iopscience.iop.org/article/10.1088/1367-2630/14/5/053037/meta}{35}],
-  [\href{https://iopscience.iop.org/article/10.1088/1367-2630/11/9/093022/meta}{37}],
-  [\href{https://www.nature.com/articles/nphys1535}{41}],
-  [\href{https://journals.aps.org/pra/abstract/10.1103/PhysRevA.87.033601}{52}],
-  and
-  [\href{https://link.springer.com/article/10.1023\%2FA\%3A1018705520999}{53}];
-  and Eqs.~(4) and (5) of
-  ref.~[\href{https://journals.aps.org/pra/abstract/10.1103/PhysRevA.90.043631}{36}])
-  that appears in an effective field theory (EFT) description of
-  low-energy scattering.  In general, an EFT description of
-  interactions would also include terms with derivatives of field
-  operators, taking the form of e.g.~Eqs.~(2) and (4) of
-  ref.~[\href{https://journals.aps.org/pra/abstract/10.1103/PhysRevA.90.043631}{36}];
-  scattering is thus generally momentum-dependent, implying the need
-  to use a momentum-dependent (commonly referred to as an
-  energy-dependent) interaction pseudo-potential.  The precise form of
-  the dominant low-energy correction to the zero-momentum contact
-  interaction in EFT is well-known (see
-  refs.~[\href{https://journals.aps.org/pra/abstract/10.1103/PhysRevA.65.043613}{55}]
-  and
-  [\href{https://journals.aps.org/pra/abstract/10.1103/PhysRevA.59.1998}{57}]
-  for studies of this correction, as well as
-  refs.~[\href{https://iopscience.iop.org/article/10.1088/1367-2630/14/5/053037/meta}{35}],
-  [\href{https://journals.aps.org/pra/abstract/10.1103/PhysRevA.90.043631}{36}],
-  and
-  [\href{https://iopscience.iop.org/article/10.1088/1367-2630/11/9/093022/meta}{37}]
-  for their application to EFT calculations), and amounts to using an
-  effective scattering length
-  \begin{align*}
-    a_{\t{eff}} = a - \f12 r_{\t{eff}} a^2 k^2,
-  \end{align*}
-  where $a$ is the zero-momentum scattering length, $r_{\t{eff}}$
-  ($\sim a$ our case) is an effective range of the leading-order
-  momentum-dependent interaction, and $k$ is the relative momentum of
-  two atoms.  We include these corrections in our all of our numerical
-  results, as discussed in the last paragraph of section III B (page
-  13, line 24):
+  \green{Note that the Hamiltonian in (5) is not the true microscopic
+    interaction Hamiltonian of AEAs, but rather a generic form for a
+    low-energy effective field theoretic description of two-body
+    interactions [17, 21, 35--37, 42--44].  There are therefore two
+    important points to keep in mind concerning our use of (5) to
+    describe two-body interactions.  First, the use of effective field
+    theory generically gives rise to divergences that must be dealt
+    with either through regularization, e.g.~of the zero-range
+    interaction potential implicitly assumed in the expression of (5)
+    [45], or through renormalization of the coupling constants in the
+    theory.  We chose the latter approach, as we will in any case find
+    it convenient to renormalize the coupling constants in the
+    effective theory developed in section III.  Second, (5) is only
+    the first term in a low-energy expansion of two-body interactions
+    in effective field theory, which generally includes additional
+    terms containing derivatives of field operators.  Derivative terms
+    correspond to the dependence of two-body scattering on the
+    relative momentum $k$ of particles involved, with $k\to0$ in the
+    zero-energy limit.  In the present case of $s$-wave scattering,
+    the leading dependence of the two-body interaction Hamiltonian on
+    the relative momentum $k$ can be captured by use of an
+    energy-dependent pseudo-potential, which amounts to using a
+    $k$-dependent effective scattering length [46].  This effective
+    scattering length can be determined by expanding the $s$-wave
+    collisional phase shift in powers of the relative momentum $k$
+    [45, 47].  Details of this expansion will depend on the
+    characteristic length scale of finite-range interactions.  In our
+    work, these corrections to (5) will be relevant only for the
+    calculation of two-body interaction energies, appearing at third
+    order in the coupling constants $G_X$.  As we are primarily
+    interested in $M$-body interactions for $M\ge3$, we defer this
+    calculation to Appendix D.  We note that our approach of using an
+    unregularized contact potential, renormalizing coupling constants,
+    and separately accounting for momentum-dependent scattering is
+    essentially the same as that used for similar calculations in
+    refs.~[35--37].}
 
-  \red{Second, the overview we provided in section II ...  neglects
-    the effect of momentum-dependent two-body scattering. ... these
-    interactions are third order in the coupling constants $G_X$.
-    ... [and therefore] will not appear in any of our three- and
-    four-body calculations ... The primary interest of our work
-    ... concerns effective $M$-body interactions for $M\ge3$; we
-    therefore defer the calculation of momentum-dependent two-body
-    interactions to Appendix D.}
-
-  A more detailed discussion of momentum-dependent corrections (and
-  thereby our use of an energy-dependent pseudo-potential) is provided
-  in Appendix D (page 36).
-
-  Given the standard use of a zero-momentum contact interaction in the
-  relevant literature, we did not consider it necessary to complicate
-  our discussion in the overview of our manuscript (section II) by
-  introducing an energy-dependent pseudo-potential.  In light of the
-  referee's comment, however, we now see that our omission of this
-  subtlety following Eq.~(2) may confuse some readers, or mislead them
-  about the nature of two-body interactions.  In order to avoid such
-  confusion without introducing major complications to our overview,
-  we have therefore appended the following sentence to the paragraph
-  containing Eq.~(5) (page 6, before line 36):
-
-  \green{Note that (2) and (5) neglect momentum-dependent corrections
-    to $s$-wave interactions, which are generally captured (in a
-    low-energy limit) by the use of an energy-dependent interaction
-    pseudo-potential [42].  These corrections will appear only at
-    third order in our effective theory, as corrections to two-body
-    interaction energies, and do not affect our discussions of
-    multi-body interactions.  We therefore defer further discussion of
-    momentum-dependent corrections to the end of section III B, and
-    their calculation to Appendix E.}
-
-  For an immediate reference on the subject matter, we have also added
-  ref.~[55] (now ref.~[42] in the revised manuscript) to the end of
-  the sentence on page 13, line 24, previously:
-
-  \red{Second, the overview we provided in section II, and
-    consequently our result in (17), neglects the effect of
-    momentum-dependent two-body scattering.}
-
-  and now:
-
-  \green{Second, the overview we provided in section II, and
-    consequently our result in (17), neglects the effect of
-    momentum-dependent two-body scattering [42].}
-
-  We hope that these additions clarify the point about our use of an
-  energy-dependent pseudo-potential.
-
-\item Concerning the second half of the previous comment by the
-  referee, on divergences and renormalization:
-
-  \blue{It is rather misleading that the authors to not point out the
-    problem at the relevant position, but only afterwards require a
-    renormalization of the divergencies without explaining the reason
-    for their appearance.}
-
-  We thank the referee for pointing out a potential point of confusion
-  in our manuscript due to the appearance of divergent terms that must
-  be regulated (in our case, by the introduction of counter-terms).
-  To address this point of confusion, we have replaced the following
-  text (page 12, line 25):
+  Expanding on our use of renormalization, we have changed the
+  following text in section III B (page 12, line 15):
 
@@
     divergences ultimately appear due to our use of effective field
-    theory to describe inter-atomic interactions in (2), rather than a
-    detailed microscopic description of two-atom scattering.
-    Divergences of this sort are a generic feature of field theories,
-    and can be dealt with using standard techniques, in particular
-    through renormalization.  We therefore renormalize our coupling
+    theory to describe inter-atomic interactions in (2) and (5),
+    rather than a detailed microscopic description of two-atom
+    scattering.  Divergences of this sort are a generic feature of
+    field theories, and can be dealt with using standard techniques
+    such as renormalization.  We therefore renormalize our coupling
     constants by introducing counter-terms $\tilde G_X$ into the
@@
 
-  We hope that this discussion helps clarify confusions regarding the
-  appearance of divergences in our effective theory.
+  We have also modified the wording in our discussion of corrections
+  to Eqs.~(2) and (5) at the end of section III B, replacing the text
+  (page 13, line 24):
+
+  \red{Second, the overview we provided in section II, and
+    consequently our result in (17), neglects the effect of
+    momentum-dependent two-body scattering. When the effective range
+    of momentum-dependent interactions is comparable to the scattering
+    lengths $a_X$, as is the case for ultracold ${}^{87}$Sr, these
+    interactions are third order in the coupling constants $G_X$. Just
+    as the $\O\p{G^2}$ counter-terms do not affect $M$-body
+    interactions until next-to-leading order (NLO), the $\O\p{G^3}$
+    momentum-dependent interactions do not come into play until next-
+    next-leading order (NNLO).}
+
+  by:
+
+  \green{Second, our result in (17) neglects the effect of
+    momentum-dependent two-body scattering.  When the effective range
+    of interactions is comparable to the scattering lengths $a_X$, as
+    is the case for ultracold ${}^{87}$Sr, these momentum-dependent
+    effects are third order in the coupling constants $G_X$.  Just as
+    the $\O\p{G^2}$ counter-terms do not affect $M$-body interactions
+    until next-to-leading order (NLO), the $\O\p{G^3}$
+    momentum-dependent terms do not come into play until
+    next-next-leading order (NNLO).}
+
+  We hope these changes and discussions help clarify our treatment of
+  two-body $s$-wave interactions, in particular addressing any
+  subtleties and potential concerns as soon as we introduce a two-body
+  interaction Hamiltonian.
 
@@
 
-  We thank the referee for pointing out a subtle point that we need to
+  We thank the referee for pointing out a subtlety that we need to
   clarify about effective theories constructed with a Schrieffer-Wolff
-  transformation.  It is true, as the referee states, that the energy
-  eigenstates $\set{\ket\alpha}$ of an interacting theory are not
-  identical to those of a non-interacting theory,
-  $\set{\ket{\alpha_0}}$, and that these sets of energy eigenstates
-  are related by a unitary transformation
-  $U=\sum_\alpha \op{\alpha_0}{\alpha}$.  While the non-interacting
-  eigenstates $\set{\ket{\alpha_0}}$ are typically simple to identify,
-  however, the interacting eigenstates $\set{\ket\alpha}$ are
-  generally difficult to represent in the same basis as the
-  Hamiltonian $H=H_0+H_{\t{int}}$ of the interacting theory, where in
-  our case $H_0$ ($H_{\t{int}}$) contains only single-body (two-body)
-  terms.  The purpose of an effective theory is then to construct an
-  {\it effective Hamiltonian} $H_{\t{eff}}$ with two key
-  properties:~(i) $H_{\t{eff}}$ is diagonal in the ``simple'' basis
-  $\set{\ket{\alpha_0}}$, and (ii) the spectrum of $H_{\t{eff}}$ is
-  identical to that of the true Hamiltonian $H$.  That is, if the
-  non-interacting and interacting Hamiltonians respectively have
-  decompositions $H_0 = \sum_\alpha E_\alpha \op{\alpha_0}$ and
-  $H = \sum_\alpha \E_\alpha \op\alpha$, an effective Hamiltonian
-  takes the form
-  $H_{\t{eff}} = U H U^\dag = \sum_\alpha \E_\alpha \op{\alpha_0}$.
-  The energy eigenstates of the effective theory are therefore
-  identical to those of the non-interacting theory, namely
-  $\set{\ket{\alpha_0}}$.
+  transformation.  Between the non-interacting Hamiltonian $H_0$, the
+  interacting Hamiltonian $H = H_0 + H_{\t{int}}$, and the effective
+  Hamiltonian $H_{\t{eff}}$ constructed from $H$ with a unitary
+  transformation, there are many different sets of energies and
+  eigenstates to keep track of.
+
+  To help clarify the relationship between $H_0$, $H$, and
+  $H_{\t{eff}}$, as well as their respective eigenstates and spectra,
+  we have re-written our opening to section III, replacing the
+  following paragraphs (page 8, line 36):
+
+  \red{As the bare two-body Hamiltonian $H_{\t{int}}$ is not diagonal
+    with respect to single-particle motional states, the problem of
+    determining many-body spectra and eigenstates nominally involves
+    all motional degrees of freedom of the atoms involved.  If we are
+    interested in addressing internal atomic states at zero
+    temperature, however, in principle we need only to consider the
+    motional ground state associated with each many-body internal
+    atomic state.  We can then ignore all excited motional states, as
+    they will be neither thermally occupied nor externally
+    interrogated.  This procedure makes the many-body problem more
+    tractable by greatly reducing the dimensionality of the relevant
+    Hilbert space.}
+
+  \red{We denote the space of all internal states of atoms in
+    single-particle, non-interacting (multi-particle, interacting)
+    motional ground states by $\H_{\t{ground}}^{\t{single}}$
+    ($\H_{\t{ground}}^{\t{multi}}$).  For sufficiently weak
+    interactions, one can construct a unitary transformation $U$
+    acting on the full many-body Hilbert space which rotates
+    $\H_{\t{ground}}^{\t{multi}}$ into $\H_{\t{ground}}^{\t{single}}$
+    [43].  The unitary $U$ can then be used to construct an {\it
+      effective Hamiltonian} $H_{\t{int}}^{\t{eff}}$ on
+    $\H_{\t{ground}}^{\t{single}}$ that reproduces the spectrum of
+    $H_{\t{int}}$ on $\H_{\t{ground}}^{\t{multi}}$, namely
+    $H_{\t{int}}^{\t{eff}}=UH_{\t{int}}U^\dag$.  This method for
+    constructing an effective theory is commonly known as the
+    Schrieffer-Wolff transformation, named after the authors of its
+    celebrated application in relating the Anderson and Kondo models
+    of magnetic impurities in metals [44].}
+
+  by:
+
+  \green{The net Hamiltonian $H = H_0 + H_{\t{int}}$ for interacting
+    AEAs on a lattice is not diagonal with respect to single-particle
+    motional state indices.  The problem of determining interacting
+    atoms' orbital excitation spectrum therefore nominally involves
+    all atomic motional degrees of freedom.  At zero temperature,
+    however, each orbital state of interacting atoms is associated
+    with a single motional ground state.  In order to compute an
+    orbital excitation spectrum at zero temperature, in principle we
+    need to identify these motional ground states.  We can then ignore
+    all excited motional states, which will be neither thermally
+    occupied nor externally interrogated.  Such a procedure would
+    drastically reduce the dimensionality of the Hilbert space
+    necessary to describe the atoms, thereby greatly simplifying any
+    description of the atoms' orbital spectrum and dynamics.  In
+    practice, however, identifying the motional ground states of
+    interacting atoms and writing down a Hamiltonian restricted to
+    this subspace is a very difficult process to carry out
+    analytically.}
+
+  \green{We denote the motional ground-state subspace of the
+    non-interacting Hamiltonian $H_0$ by
+    $\H_{\t{ground}}^{\t{single}}$, and the motional ground-state
+    subspace of the interacting Hamiltonian $H = H_0 + H_{\t{int}}$ by
+    $\H_{\t{ground}}^{\t{multi}}$.  That is, all atomic wavefunctions
+    for states within $\H_{\t{ground}}^{\t{single}}$ are described by
+    $\phi_{i,0}$, while the atomic wavefunctions for states within
+    $\H_{\t{ground}}^{\t{multi}}$ are generally unknown, and are in
+    principle determined by minimizing the energy of a state with
+    respect to its motional degrees of freedom.  Both
+    $\H_{\t{ground}}^{\t{single}}$ and $\H_{\t{ground}}^{\t{multi}}$
+    are subspaces of the full Hilbert space $\H_{\t{full}}$.  When
+    interactions are sufficiently weak compared to the spectral gap
+    $\Delta$ of the non-interacting Hamiltonian $H_0$ between
+    $\H_{\t{ground}}^{\t{single}}$ and its orthogonal complement
+    $\H_{\t{full}}\setminus\H_{\t{ground}}^{\t{single}}$, one can
+    identify a particular unitary operator $U$ (acting on the full
+    Hilbert space $\H_{\t{full}}$) which rotates
+    $\H_{\t{ground}}^{\t{multi}}$ into $\H_{\t{ground}}^{\t{single}}$
+    [49].  This unitary $U$ can be used to construct an {\it effective
+      Hamiltonian} $H_{\t{eff}} = U H U^\dag$ with two key properties:
+    (i) $H_{\t{eff}}$ is diagonal in the same (known) basis as the
+    non-interacting Hamiltonian $H_0$, and (ii) the spectrum of
+    $H_{\t{eff}}$ on $\H_{\t{ground}}^{\t{single}}$ is identical to
+    that of the interacting Hamiltonian $H$ on
+    $\H_{\t{ground}}^{\t{multi}}$.  The use of an effective
+    Hamiltonian $H_{\t{eff}}$ thus side-steps the need to identify
+    $\H_{\t{ground}}^{\t{multi}}$ in order to compute the orbital
+    spectrum of $H$ at zero temperature.  This method for constructing
+    an effective theory is commonly known as the Schrieffer-Wolff
+    transformation, named after the authors of its celebrated
+    application in relating the Anderson and Kondo models of magnetic
+    impurities in metals [50].}
+
+  We have also modified the text pointed out by the referee, changing
+  (page 10, line 19):
+
+  \red{Finally, the effective theory involves no corrections to the
+    many-body energy eigenstates.  Including such corrections would
+    invalidate the effective theory, which preserves the eigenstates
+    of the non-interacting Hamiltonian $H_0$ on
+    $\H_{\t{ground}}^{\t{single}}$.}
+
+  to:
+
+  \green{Finally, the effective theory involves no corrections to the
+    non-interacting many-body energy eigenstates; the purpose of the
+    effective Hamiltonian $H_{\t{eff}} = H_0 + H_{\t{int}}^{\t{eff}}$
+    is to reproduce, on the eigenstates of the non-interacting
+    Hamiltonian $H_0$ within $\H_{\t{ground}}^{\t{single}}$, the
+    spectrum of the interacting Hamiltonian $H = H_0 + H_{\t{int}}$ on
+    $\H_{\t{ground}}^{\t{multi}}$.  Modifying the eigenstates of the
+    non-interacting Hamiltonian $H_0$ on
+    $\H_{\t{ground}}^{\t{single}}$ thus invalidates the effective
+    theory.}
+
+  We hope that these changes clarify subtleties of the
+  Schrieffer-Wolff transformation, and in particular the fact that the
+  effective Hamiltonian $H_{\t{eff}}$ has the spectrum of the
+  interacting Hamiltonian $H$ on the eigenstates of the
+  non-interacting Hamiltonian $H_0$.
+
+\item Concerning:
+
+  \blue{In addition, the is also an important scientific question. The
+    authors use renormalization of the theory by adding counter terms
+    to remove the diverging contributions. This approach might provide
+    the correct result however a microscopic derivation and motivation
+    based on the pseudo-potential would be certainly help to justify
+    this approach.}
 

### 2019-02-07 multi_body_interactions/papers/theory/paper.tex
updates/changes from responses to NJP referee 1

@@
 % general physics / math packages and commands
-\usepackage{physics,amsmath,amssymb,braket,dsfont,bm}
+\usepackage{physics,amsmath,amssymb,braket,dsfont}
 \renewcommand{\t}{\text} % text in math mode
@@
 
+\renewcommand{\d}{\text{d}}
 \newcommand{\g}{\text{g}}
@@
   the prospect of performing quantum simulations of spin models and
-  lattice field theories exhibiting SU($N$) symmetry.  In order to
-  analyze recent experiments preparing high density samples of
+  lattice field theories exhibiting SU($N$) symmetry.  Motivated in
+  particular by recent experiments preparing high density samples of
   strongly interacting ${}^{87}$Sr atoms in a three-dimensional
@@
   interactions, where $N$ is the number of atomic nuclear spin levels.
-  We fully characterize the low-lying excitations in our effective
-  theory, and compare predictions of many-body interaction energies
-  with direct measurements of many-body excitation spectra in an
-  optical lattice clock.  Our work makes the first step in enabling a
-  controlled, bottom-up experimental investigation of multi-body
-  SU($N$) physics.
+  Our theory is limited to the experimental regime of
+  \begin{enumerate*}
+  \item a deep lattice, with
+  \item at most one atom occupying each nuclear spin state on any
+    lattice site,
+  \end{enumerate*}
+  where the latter restriction is a consequence of initial
+  ground-state preparation.  We fully characterize the low-lying
+  excitations in our effective theory, and compare predictions of
+  many-body interaction energies with direct measurements of many-body
+  excitation spectra in an optical lattice clock.  Our work makes the
+  first step in enabling a controlled, bottom-up experimental
+  investigation of multi-body SU($N$) physics.
 \end{abstract}
@@
 thereby enabling experimental studies of multi-body SU($N$) physics
-through the exquisite capabilities with OLCs.  We briefly summarize
-the experimental procedures relevant to our work in Appendix
-\ref{sec:experiment_summary}.
+through the exquisite capabilities with OLCs.  Our theory is limited
+to the experimental regime of at most one atom occupying each nuclear
+spin state on any lattice site, which is a consequence initial
+ground-state preparation.
 
@@
 for the first time with fermions that have internal degrees of freedom
-and multiple collisional parameters.  Nonetheless, due to the SU($N$)
-symmetry of these collisions, we are able to find a simple way to
-express effective multi-body Hamiltonians and fully characterize their
-low-lying many-body energy eigenstates.  While some past work has
-detected experimental signatures of multi-body interactions in the
-form of quantum phase revivals \cite{will2010timeresolved}, we perform
-a direct comparison of the many-body interaction energies predicted by
-our low-energy effective theory with experimental measurements of
-density-dependent atomic energy spectra performed in
+and multiple collisional parameters.  We perform a direct comparison
+between the many-body interaction energies predicted by our low-energy
+effective theory and the experimental measurements of the
+density-dependent orbital excitation spectra performed in
 ref.~\cite{goban2018emergence}, similarly to the measurements with
-bosons performed in ref.~\cite{mark2011precision}.
+bosons performed in ref.~\cite{mark2011precision}.  To facilitate this
+comparison of excitation spectra and characterize the low-lying
+excitations in our effective theory, we consider a restriction of the
+theory to states with at most one orbital excitation per lattice site,
+and find that the SU($N$) symmetry of atomic collisions allow
+effective multi-body interactions to take a remarkably simple form.
+We note that past work has also detected experimental signatures of
+multi-body interactions in the form of quantum phase revivals
+\cite{will2010timeresolved}.
 
 The remainder of this paper is structured as follows.  In section
-\ref{sec:overview} we provide an overview of the relevant one- and
-two-body physics of ultracold atoms on a lattice, and summarize our
-main result of effective multi-body interactions between these atoms.
+\ref{sec:overview} we summarize the experimental procedures relevant
+to our work, provide an overview one- and two-body physics of
+ultracold atoms on a lattice, and preview our main technical results.
 In section \ref{sec:low_energy} we discuss our method for deriving a
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Overview}
+\section{Background and overview}
 \label{sec:overview}
 
-Consider a collection of fermionic atoms with total nuclear spin $I$
-and two atomic orbital states loaded into a translationally invariant
-3-D optical lattice.  While an external trapping potential will
-generally break discrete translational symmetry of the lattice, any
-background inhomogeneity can be made negligible by spectroscopically
-addressing a sufficiently small region of the lattice
-\cite{goban2018emergence}.  Throughout this paper, we work strictly in
-the deep-lattice regime with negligible tunneling between lattice
-sites.  We also assume a so-called ``magic-wavelength'' lattice for
-which both atomic orbital states experience identical lattice
-potentials \cite{ye2008quantum}.  The single-particle Hamiltonian of
-the atoms can then be written in the form
+The work in this paper is closely tied to the experimental work in
+ref.~\cite{goban2018emergence}; we begin with a short summary of the
+relevant experimental procedures therein.  The experiment begins by
+preparing a degenerate gas of $10^4$-$10^5$ (fermionic) ${}^{87}$Sr
+atoms in a uniform mixture of their ten nuclear spin states and at
+$\sim0.1$ of their fermi temperature ($\sim10$ nanokelvin)
+\cite{campbell2017fermidegenerate, marti2018imaging}.  This gas is
+loaded into a primitive cubic optical lattice at the ``magic
+wavelength'' for which both ground (${}^1S_0$) and first-excited
+(${}^3P_0$) electronic orbital states of the atoms experience the same
+lattice potential \cite{ye2008quantum}.  Lattice depths along the
+principal axes of the lattice are roughly equal in magnitude, with a
+geometric mean that can be varied from 30 to 80 $E_{\t{R}}$, where
+$E_{\t{R}}\approx3.5\times2\pi~\t{kHz}$ is the lattice photon recoil
+energy of the atoms (with the reduced Planck constant $\hbar=1$
+throughout this paper).  These lattice depths are sufficiently large
+as to neglect tunneling on the time scales relevant to the experiment.
+The temperature of the atoms is also low enough to neglect thermal
+occupation of states outside the ground-state manifold.
+
+Once loaded into an optical lattice lattice, atoms are addressed by an
+external (``clock'') interrogation laser with an ultranarrow (26 mHz)
+linewidth, detuned by $\Delta$ from the single-atom ${}^1S_0-{}^3P_0$
+transition frequency $\omega_0$.  After a fixed interrogation time,
+the experiment turns off the interrogation laser, removes all
+ground-state (${}^1S_0$) atoms from the lattice, and uses absorption
+imaging to count any remaining excited-state (${}^3P_0$) atoms.
+Non-interacting atoms in singly-occupied lattice sites undergo Rabi
+oscillations when $\Delta=0$ during the interrogation time.  The
+orbital excitation spectrum of multiply-occupied lattice sites,
+however, is shifted by inter-atomic interactions, which results in
+spectroscopic peaks (i.e.~local maxima in excited-state atom counts)
+away from $\Delta=0$.  A sweep across different detunings $\Delta$ (on
+the scale of inter-atomic interaction energies) thus constitutes a
+measurement of the many-body orbital excitation spectrum.  We note
+that this spectroscopic protocol addresses only singly-excited orbital
+states of any lattice site.  Doubly-excited states are manifestly
+off-resonant for single-photon absorption processes as
+$\abs{\Delta}\ll\omega_0$, while multi-photon absorption processes are
+always off-resonant due to
+\begin{enumerate*}
+\item the non-linearity ($\sim$kHz) of orbital excitation energies for
+  a collection of interacting atoms, and
+\item the ultranarrow linewidth ($\sim$mHz) of the interrogation
+  laser.
+\end{enumerate*}
+
+Consider now a collection of fermionic atoms with total nuclear spin
+$I$ and two atomic orbital states loaded into a translationally
+invariant 3-D optical lattice.  While an external trapping potential
+will generally break discrete translational symmetry of the lattice,
+any background inhomogeneity can be made negligible by
+spectroscopically addressing a sufficiently small region of the
+lattice \cite{goban2018emergence}.  Throughout this paper, we work
+strictly in the deep-lattice regime with negligible tunneling between
+lattice sites, and assume that both atomic orbital states experience
+identical lattice potentials.  The single-particle Hamiltonian of the
+atoms can then be written in the form
 \begin{align}
@@
 $E_n=\p{3/2+n_\x+n_\y+n_\z}\omega$ for an on-site angular trap
-frequency $\omega$ (with the reduced Planck constant $\hbar=1$
-throughout this paper), but in general the aharmonicity of the lattice
+frequency $\omega$, but in general the aharmonicity of the lattice
 potential will cause a non-negligible shift in motional state
@@
   H_{\t{int}} = \sum_{\substack{\mu<\nu\\s}}
-  G_s \int d^3x~ \hat\rho_{\mu s} \hat\rho_{\nu s}
+  G_s \int \d^3x~ \hat\rho_{\mu s} \hat\rho_{\nu s}
   + G_+ \sum_{\mu,\nu}
-  \int d^3x~ \hat\rho_{\mu,\e} \hat\rho_{\nu,\g}
-  + G_- \sum_{\mu,\nu} \int d^3x~
+  \int \d^3x~ \hat\rho_{\mu,\e} \hat\rho_{\nu,\g}
+  + G_- \sum_{\mu,\nu} \int \d^3x~
   \hat\psi_{\mu,\e}^\dag \hat\psi_{\nu,\g}^\dag
@@
   H_{\t{int}}
-  = \f12 \sum_{\substack{q,r,s,t\\\mu,\nu}} G^{qr}_{st} \int d^3x~
+  = \f12 \sum_{\substack{q,r,s,t\\\mu,\nu}} G^{qr}_{st} \int \d^3x~
   \hat\psi_{\mu s}^\dag \hat\psi_{\nu t}^\dag
@@
 coupling constant between the two-atom states
-$(\mu,q)+(\nu,r)\leftrightarrow(\mu,s)+(\nu,t)$.  Note that
-\eqref{eq:H_int_start} and \eqref{eq:H_int_fields} neglect
-momentum-dependent corrections to $s$-wave interactions, which are
-generally captured (in a low-energy limit) by the use of an
-energy-dependent interaction pseudo-potential \cite{blume2002fermi}.
-These corrections will appear only at third order in our effective
-theory, as corrections to two-body interaction energies, and do not
-affect our discussions of multi-body interactions.  We therefore defer
-further discussion of momentum-dependent corrections to the end of
-section \ref{sec:two_body}, and their calculation to Appendix
-\ref{sec:momentum_dependence}.
+$(\mu,q)+(\nu,r)\leftrightarrow(\mu,s)+(\nu,t)$.
+
+Note that the Hamiltonian in \eqref{eq:H_int_fields} is not the true
+microscopic interaction Hamiltonian of AEAs, but rather a generic form
+for a low-energy effective field theoretic description of two-body
+interactions \cite{scazza2014observation, cazalilla2009ultracold,
+  johnson2012effective, yin2014universal, johnson2009effective,
+  gorshkov2010twoorbital, wall2013strongly, busch1998two}.  There are
+therefore two important points to keep in mind concerning our use of
+\eqref{eq:H_int_fields} to describe two-body interactions.  First, the
+use of effective field theory generically gives rise to divergences
+that must be dealt with either through regularization, e.g.~of the
+zero-range interaction potential implicitly assumed in the expression
+of \eqref{eq:H_int_fields} \cite{giorgini2008theory}, or through
+renormalization of the coupling constants in the theory.  We chose the
+latter approach, as we will in any case find it convenient to
+renormalize the coupling constants in the effective theory developed
+in section \ref{sec:low_energy}.  Second, \eqref{eq:H_int_fields} is
+only the first term in a low-energy expansion of two-body interactions
+in effective field theory, which generally includes additional terms
+containing derivatives of field operators.  Derivative terms
+correspond to the dependence of two-body scattering on the relative
+momentum $k$ of particles involved, with $k\to0$ in the zero-energy
+limit.  In the present case of $s$-wave scattering, the leading
+dependence of the two-body interaction Hamiltonian on the relative
+momentum $k$ can be captured by use of an energy-dependent
+pseudo-potential, which amounts to using a $k$-dependent effective
+scattering length \cite{blume2002fermi}.  This effective scattering
+length can be determined by expanding the $s$-wave collisional phase
+shift in powers of the relative momentum $k$ \cite{giorgini2008theory,
+  flambaum1999analytical}.  Details of this expansion will depend on
+the characteristic length scale of finite-range interactions.  In our
+work, these corrections to \eqref{eq:H_int_fields} will be relevant
+only for the calculation of two-body interaction energies, appearing
+at third order in the coupling constants $G_X$.  As we are primarily
+interested in $M$-body interactions for $M\ge3$, we defer this
+calculation to Appendix \ref{sec:momentum_dependence}.  We note that
+our approach of using an unregularized contact potential,
+renormalizing coupling constants, and separately accounting for
+momentum-dependent scattering is essentially the same as that used for
+similar calculations in refs.~\cite{johnson2012effective,
+  yin2014universal, johnson2009effective}.
 
@@
   K^{k\ell}_{mn}
-  \equiv \int d^3x~ \phi_{im}^* \phi_{in}^* \phi_{i\ell} \phi_{ik},
+  \equiv \int \d^3x~ \phi_{im}^* \phi_{in}^* \phi_{i\ell} \phi_{ik},
   \label{eq:K_klmn}
@@
 coherently address states with a single orbital excitation per lattice
-site (see Appendix \ref{sec:experiment_summary})
-\cite{goban2018emergence}.  The temperature of the atoms in these
-experiments is much smaller than single-atom motional excitation
-energies $\Delta_n\equiv E_n-E_0$ for $n>0$, which implies a
-negligible thermal occupation of excited motional states.  A naive
-calculation of the multi-body orbital excitation spectrum, however,
-assuming that atoms only occupy motional ground states with spatial
-wavefunctions $\phi_{i,0}$, fails to reproduce the experimentally
-observed orbital excitation spectrum of multiply-occupied lattice
-sites.  The reason that such a calculation fails to reproduce the
-observed excitation spectrum has to do with the fact that the two-body
-Hamiltonian $H_{\t{int}}$ in \eqref{eq:H_int} is not diagonal in the
-basis of single-particle motional eigenstates (i.e.~$\phi_{in}$).  For
-interacting atoms, the true motional ground states are therefore
-admixtures of the non-interacting motional eigenstates; when the
-off-diagonal elements of $H_{\t{int}}$ are comparable in magnitude to
-the excitation energies $\Delta_n$, this admixture can have
-non-negligible weight outside the non-interacting motional
-ground-state subspace.  In order to compute the correct many-body
-orbital excitation spectrum, we develop a low-energy effective theory
-for such atoms.  As our work is motivated by experiments that
-initialize all atoms in identical (i.e.~ground) orbital and motional
-states, to simplify our theory we assume that any $N$ atoms on a
-single lattice site occupy $N$ nuclear spin states.  Ground-state
-preparation implies that multiple occupation of a single nuclear spin
-state on any given lattice site is initially forbidden by fermionic
-statistics, and subsequently cannot occur in the absence of inter-site
-effects or hyperfine coupling.
+site \cite{goban2018emergence}.  At ultracold temperatures well below
+the single-atom motional excitation energies $\Delta_n\equiv E_n-E_0$
+for $n>0$, atoms only occupy their motional ground state in the
+lattice.  For this reason, it is common to map the description of
+these atoms onto a single-band Hubbard model that captures all
+dynamics within the subspace of motional ground states, i.e.~with
+single-atom wavefunctions $\phi_{i,0}$.  When enough atoms occupy the
+same lattice site, however, their interaction energy can approach the
+scale of the motional excitation energies $\Delta_n$.  In this case,
+normally far-off-resonant terms in the interaction Hamiltonian of
+\eqref{eq:H_int} that create atoms in excited motional states,
+e.g.~$\sim\c_{n\mu s}^\dag \c_{0,\nu t}^\dag \c_{0,\nu r} \c_{0,\mu
+  q}$ with $n>0$, become relevant.  The true motional ground state of
+a collection of interacting atoms is then an admixture of the
+non-interacting motional eigenstates, and a naive Hubbard model that
+assumes atomic wavefunctions $\phi_{i,0}$ will fail to reproduce the
+interacting atoms' orbital excitation spectrum.
+
+In order to understand interacting atoms' orbital excitation spectrum
+in the context of a zero-temperature Hubbard-type model that
+eliminates non-dynamical (i.e.~ground-state) motional degrees of
+freedom, we develop a low-energy effective theory of interacting AEAs
+in a deep lattice.  To simplify our theory, we assume that any $N$
+atoms on a single lattice site occupy $N$ nuclear spin states.
+Ground-state preparation of the atoms in an experiment implies that
+multiple occupation of a single nuclear spin state on any given
+lattice site is initially forbidden by fermionic statistics, and
+subsequently cannot occur in the absence of inter-site effects or
+hyperfine coupling between nuclear spin states.
 
@@
 where $H_M^{(p)}$ is an $M$-body Hamiltonian of order $p$ in the
-coupling constants $G^{qr}_{st}$, $I$ is the total nuclear spin of
-each atom (e.g.~$I=9/2$ for ${}^{87}$Sr), and the sum terminates at
-$2I+1$ because this is the largest number of atoms which may initially
-occupy a single lattice site.  We explicitly compute all $M$-body
+coupling constants $G_X$, $I$ is the total nuclear spin of each atom
+(e.g.~$I=9/2$ for ${}^{87}$Sr), and the sum terminates at $2I+1$
+because this is the largest number of atoms which may initially occupy
+a single lattice site.  We explicitly compute all $M$-body
 Hamiltonians $H_M\equiv\sum_p H_M^{(p)}$ through order $p=3$, yielding
-effective two-, three-, and four-body interactions.  To facilitate
-understanding and comparison with experimental measurements of
-density-dependent atomic energy spectra [34], in which states with
-multiple orbital excitations on a single lattice site are
-spectroscopically off-resonant, we also consider a restriction of our
-theory to states with at most one orbital excitation per lattice site.
-In this case, we find that SU($N$) symmetry allows us to express
-multi-body Hamiltonians in the simple form
+effective two-, three-, and four-body interactions.  To
+\begin{enumerate*}
+\item facilitate a comparison with the experimental measurements of
+  many-body orbital excitation spectra performed in
+  ref.~\cite{goban2018emergence} and
+\item characterize the low-lying excitations in our effective theory,
+\end{enumerate*}
+we also consider a restriction of the multi-body Hamiltonians $H_M$ to
+states with at most one orbital excitation per lattice site.  Under
+this restriction, we find that the SU($N$) symmetry of atomic
+collisions allows us to express multi-body Hamiltonians in the simple
+form
 \begin{align}
   H_M = \sum_{\abs{\set{\mu_j}}=M}
-  \p{U_{M,\g} \prod_{\alpha=1}^M \n_{\mu_\alpha,\g}
-    + U_{M,+} \n_{\mu_1,\e} \prod_{\alpha=2}^M \n_{\mu_\alpha,\g}
-    + U_{M,-} \c_{\mu_1,\g}^\dag \c_{\mu_2,\e}^\dag
-    \c_{\mu_2,\g} \c_{\mu_1,\e} \prod_{\alpha=3}^M \n_{\mu_\alpha,\g}},
+  H_2^{(\mu_1,\mu_2)} \prod_{\alpha=3}^M \n_{\mu_\alpha,\g},
   \label{eq:H_M_preview}
 \end{align}
-where $\c_{\mu s}\equiv\c_{0,\mu s}$ is a fermionic annihilation
-operator for an atom in the motional ground state with nuclear spin
-$\mu$ and orbital state $s$;
-$\n_{\mu s}\equiv \c_{\mu s}^\dag \c_{\mu s}$ is a number operator;
-and the sum is performed over all choices of nuclear spins $\mu_j$
-with $j=1,2,\cdots,M$ for which all $\mu_j$ are distinct, or
-equivalently all choices of $\mu_j$ for which the set $\set{\mu_j}$
-contains $M$ elements, for a total of ${2I+1\choose M}\times\p{M!}$
-nuclear spin combinations.  The coefficients $U_X$ can be expressed in
-terms of the coupling constants of the bare two-body interaction
-Hamiltonian (see Appendix \ref{sec:U_X}).  The key feature of the
-$M$-body interactions in \eqref{eq:H_M_preview} is that they take a
-form identical to the two-body interactions between motional
-ground-state atoms, but with the addition of $M-2$ spectator atoms.
+where $H_2^{(\mu_1,\mu_2)}$ is a two-body Hamiltonian addressing atoms
+with nuclear spin $\mu_1,\mu_2$;
+$\n_{\mu s}=\c_{\mu s}^\dag\c_{\mu s}$ is a number operator for atoms
+with nuclear state $\mu$ and orbital state $s$; and the sum is
+performed over all choices of nuclear spins $\mu_j$ with
+$j=1,2,\cdots,M$ for which all $\mu_j$ are distinct, or equivalently
+all choices of $\mu_j$ for which the set $\set{\mu_j}$ contains $M$
+elements, for a total of ${2I+1\choose M}\times\p{M!}$ nuclear spin
+combinations.  The key feature of the $M$-body interactions in
+\eqref{eq:H_M_preview} is that they ultimately take the same form as
+two-body interactions, but with the addition of $M-2$ spectator atoms.
 This form is a direct consequence of the SU($N$) symmetry of
@@
 
+\newpage
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-As the bare two-body Hamiltonian $H_{\t{int}}$ is not diagonal with
-respect to single-particle motional states, the problem of determining
-many-body spectra and eigenstates nominally involves all motional
-degrees of freedom of the atoms involved.  If we are interested in
-addressing internal atomic states at zero temperature, however, in
-principle we need only to consider the motional ground state
-associated with each many-body internal atomic state.  We can then
-ignore all excited motional states, as they will be neither thermally
-occupied nor externally interrogated.  This procedure makes the
-many-body problem more tractable by greatly reducing the
-dimensionality of the relevant Hilbert space.
-
-We denote the space of all internal states of atoms in
-single-particle, non-interacting (multi-particle, interacting)
-motional ground states by $\H_{\t{ground}}^{\t{single}}$
-($\H_{\t{ground}}^{\t{multi}}$).  For sufficiently weak interactions,
-one can construct a unitary transformation $U$ acting on the full
-many-body Hilbert space which rotates $\H_{\t{ground}}^{\t{multi}}$
-into $\H_{\t{ground}}^{\t{single}}$ \cite{bravyi2011schrieffer}.  The
-unitary $U$ can then be used to construct an {\it effective
-  Hamiltonian} $H_{\t{int}}^{\t{eff}}$ on
-$\H_{\t{ground}}^{\t{single}}$ that reproduces the spectrum of
-$H_{\t{int}}$ on $\H_{\t{ground}}^{\t{multi}}$, namely
-$H_{\t{int}}^{\t{eff}}=UH_{\t{int}}U^\dag$.  This method for
+The net Hamiltonian $H = H_0 + H_{\t{int}}$ for interacting AEAs on a
+lattice is not diagonal with respect to single-particle motional state
+indices.  The problem of determining interacting atoms' orbital
+excitation spectrum therefore nominally involves all atomic motional
+degrees of freedom.  At zero temperature, however, each orbital state
+of interacting atoms is associated with a single motional ground
+state.  In order to compute an orbital excitation spectrum at zero
+temperature, in principle we need to identify these motional ground
+states.  We can then ignore all excited motional states, which will be
+neither thermally occupied nor externally interrogated.  Such a
+procedure would drastically reduce the dimensionality of the Hilbert
+space necessary to describe the atoms, thereby greatly simplifying any
+description of the atoms' orbital spectrum and dynamics.  In practice,
+however, identifying the motional ground states of interacting atoms
+and writing down a Hamiltonian restricted to this subspace is a very
+difficult process to carry out analytically.
+
+We denote the motional ground-state subspace of the non-interacting
+Hamiltonian $H_0$ by $\H_{\t{ground}}^{\t{single}}$, and the motional
+ground-state subspace of the interacting Hamiltonian
+$H = H_0 + H_{\t{int}}$ by $\H_{\t{ground}}^{\t{multi}}$.  That is,
+all atomic wavefunctions for states within
+$\H_{\t{ground}}^{\t{single}}$ are described by $\phi_{i,0}$, while
+the atomic wavefunctions for states within
+$\H_{\t{ground}}^{\t{multi}}$ are generally unknown, and are in
+principle determined by minimizing the energy of a state with respect
+to its motional degrees of freedom.  Both
+$\H_{\t{ground}}^{\t{single}}$ and $\H_{\t{ground}}^{\t{multi}}$ are
+subspaces of the full Hilbert space $\H_{\t{full}}$.  When
+interactions are sufficiently weak compared to the spectral gap
+$\Delta$ of the non-interacting Hamiltonian $H_0$ between
+$\H_{\t{ground}}^{\t{single}}$ and its orthogonal complement
+$\H_{\t{full}}\setminus\H_{\t{ground}}^{\t{single}}$, one can identify
+a particular unitary operator $U$ (acting on the full Hilbert space
+$\H_{\t{full}}$) which rotates $\H_{\t{ground}}^{\t{multi}}$ into
+$\H_{\t{ground}}^{\t{single}}$ \cite{bravyi2011schrieffer}.  This
+unitary $U$ can be used to construct an {\it effective Hamiltonian}
+$H_{\t{eff}} = U H U^\dag$ with two key properties:
+\begin{enumerate*}
+\item $H_{\t{eff}}$ is diagonal in the same (known) basis as the
+  non-interacting Hamiltonian $H_0$, and
+\item the spectrum of $H_{\t{eff}}$ on $\H_{\t{ground}}^{\t{single}}$
+  is identical to that of the interacting Hamiltonian $H$ on
+  $\H_{\t{ground}}^{\t{multi}}$.
+\end{enumerate*}
+The use of an effective Hamiltonian $H_{\t{eff}}$ thus side-steps the
+need to identify $\H_{\t{ground}}^{\t{multi}}$ in order to compute the
+orbital spectrum of $H$ at zero temperature.  This method for
 constructing an effective theory is commonly known as the
@@
 (i.e.~interacting) and unperturbed (i.e.~non-interacting) Hamiltonian,
-we derive an expansion for the effective interaction Hamiltonian
+we derive an expansion for an effective interaction Hamiltonian
 $H_{\t{int}}^{\t{eff}}$ in terms two-body interaction Hamiltonian
@@
 where $H_{\t{int}}^{(p)}$ is order $p$ in $H_{\t{int}}$.  Letting
-$\E_0$ denote the orthogonal complement of
-$\H_{\t{ground}}^{\t{single}}$ (i.e.~$\E_0$ is the space of all states
-with at least one atom in an excited motional state), $\B_0\p{\E_0}$
-denote an eigenbasis of $\E_0$ with respect to the single-particle
-Hamiltonian $H_0$, and $E_\alpha$ denote the motional energy (with
-respect to $H_0$) of a state $\ket\alpha\in\E_0$ relative to the
-corresponding motional ground-state energy, we define the operator
+$\E_0\equiv \H_{\t{full}}\setminus\H_{\t{ground}}^{\t{single}}$ denote
+the orthogonal complement of $\H_{\t{ground}}^{\t{single}}$
+(i.e.~$\E_0$ is the space of all states with at least one atom in an
+excited motional state), $\B_0\p{\E_0}$ denote an eigenbasis of $\E_0$
+with respect to the single-particle Hamiltonian $H_0$, and $E_\alpha$
+denote the motional energy (with respect to $H_0$) of a state
+$\ket\alpha\in\E_0$ relative to the corresponding motional
+ground-state energy, we define the operator
 \begin{align}
@@
 \end{align}
-where $\sp{X,Y}_+\equiv XY+YX$.  The naive idea of neglecting all
-excited states and using $H_{\t{int}}$ directly to describe the
-spectrum of single-particle motional ground states is thus equivalent
-to truncating our expansion for $H_{\t{int}}^{\t{eff}}$ at first
-order.  In addition to this first order term, the expansion involves
-{\it effective corrections} to the action of $H_{\t{int}}$ on the
-single-particle motional ground states (i.e.~on
-$\H_{\t{ground}}^{\t{single}}$) in the form of higher-order terms with
-{\it intermediate} or {\it virtual occupation} of excited states,
-manifest in $\I$.
+where $\sp{X,Y}_+\equiv XY+YX$.  Writing down a single-band Hubbard
+model that simply neglects excited atomic motional states and uses
+$H_{\t{int}}$ directly to describe the orbital spectrum of interacting
+atoms is thus equivalent to truncating our expansion for
+$H_{\t{int}}^{\t{eff}}$ at first order.  In addition to this first
+order term, the expansion involves {\it effective corrections} to the
+action of $H_{\t{int}}$ on the single-particle motional ground states
+(i.e.~on $\H_{\t{ground}}^{\t{single}}$) in the form of higher-order
+terms with {\it intermediate} or {\it virtual occupation} of excited
+states, manifest in $\I$.
 
@@
   bravyi2011schrieffer}.  Finally, the effective theory involves no
-corrections to the many-body energy eigenstates.  Including such
-corrections would invalidate the effective theory, which preserves the
-eigenstates of the non-interacting Hamiltonian $H_0$ on
-$\H_{\t{ground}}^{\t{single}}$.
+corrections to the non-interacting many-body energy eigenstates; the
+purpose of the effective Hamiltonian
+$H_{\t{eff}} = H_0 + H_{\t{int}}^{\t{eff}}$ is to reproduce, on the
+eigenstates of the non-interacting Hamiltonian $H_0$ within
+$\H_{\t{ground}}^{\t{single}}$, the spectrum of the interacting
+Hamiltonian $H = H_0 + H_{\t{int}}$ on $\H_{\t{ground}}^{\t{multi}}$.
+Modifying the eigenstates of the non-interacting Hamiltonian $H_0$ on
+$\H_{\t{ground}}^{\t{single}}$ thus invalidates the effective theory.
 
@@
 appear due to our use of effective field theory to describe
-inter-atomic interactions in \eqref{eq:H_int_start}, rather than a
-detailed microscopic description of two-atom scattering.  Divergences
-of this sort are a generic feature of field theories, and can be dealt
-with using standard techniques, in particular through renormalization.
-We therefore renormalize our coupling constants by introducing
-counter-terms $\tilde G_X$ into the effective interaction Hamiltonians
-via $G_X\to G_X+\tilde G_X$.  The values of counter-terms can be fixed
-by enforcing that $G_X$ is equal to the net two-body interaction
-strength, which has the added benefit of allowing us to express the
-effective two-body interactions in terms of net, physically observable
-two-body coupling constants on the right-hand side of
-\eqref{eq:H_2_expansion}, rather than in terms of long sums at all
-orders of the bare coupling constants.  Diagrammatically, this
-renormalization condition is
+inter-atomic interactions in \eqref{eq:H_int_start} and
+\eqref{eq:H_int_fields}, rather than a detailed microscopic
+description of two-atom scattering.  Divergences of this sort are a
+generic feature of field theories, and can be dealt with using
+standard techniques such as renormalization.  We therefore renormalize
+our coupling constants by introducing counter-terms $\tilde G_X$ into
+the effective interaction Hamiltonians via $G_X\to G_X+\tilde G_X$.
+The values of counter-terms can be fixed by enforcing that $G_X$ is
+equal to the net two-body interaction strength, which has the added
+benefit of allowing us to express the effective two-body interactions
+in terms of net, physically observable two-body coupling constants on
+the right-hand side of \eqref{eq:H_2_expansion}, rather than in terms
+of long sums at all orders of the bare coupling constants.
+Diagrammatically, this renormalization condition is
 \begin{align}
@@
 
-Second, the overview we provided in section \ref{sec:overview}, and
-consequently our result in \eqref{eq:H_2}, neglects the effect of
-momentum-dependent two-body scattering \cite{blume2002fermi}.  When
-the effective range of momentum-dependent interactions is comparable
-to the scattering lengths $a_X$, as is the case for ultracold
-${}^{87}$Sr, these interactions are third order in the coupling
-constants $G_X$.  Just as the $\O\p{G^2}$ counter-terms do not affect
-$M$-body interactions until next-to-leading order (NLO), the
-$\O\p{G^3}$ momentum-dependent interactions do not come into play
-until next-next-leading order (NNLO).  Given that we develop our
-low-energy effective theory through third order in the coupling
-constants, these interactions will not appear in any of our three- and
-four-body calculations, but they do need to be considered in the
-calculation of pair-wise interaction energies.  The primary interest
-of our work, however, concerns effective $M$-body interactions for
-$M\ge3$; we therefore defer the calculation of momentum-dependent
-two-body interactions to Appendix \ref{sec:momentum_dependence}.
+Second, our result in \eqref{eq:H_2} neglects the effect of
+momentum-dependent two-body scattering.  When the effective range of
+interactions is comparable to the scattering lengths $a_X$, as is the
+case for ultracold ${}^{87}$Sr, these momentum-dependent effects are
+third order in the coupling constants $G_X$.  Just as the $\O\p{G^2}$
+counter-terms do not affect $M$-body interactions until
+next-to-leading order (NLO), the $\O\p{G^3}$ momentum-dependent terms
+do not come into play until next-next-leading order (NNLO).  Given
+that we develop our low-energy effective theory through third order in
+the coupling constants, these interactions will not appear in any of
+our three- and four-body calculations, but they do need to be
+considered in the calculation of pair-wise interaction energies.  The
+primary interest of our work, however, concerns effective $M$-body
+interactions for $M\ge3$; we therefore defer the calculation of
+momentum-dependent two-body interactions to Appendix
+\ref{sec:momentum_dependence}.
 
@@
 Current experiments with ultracold ${}^{87}$Sr on a lattice can
-coherently address ground states and their low-lying orbital
-excitations for up to five atoms per lattice site
-\cite{goban2018emergence} (see Appendix \ref{sec:experiment_summary}
-for a summary of relevant experimental procedures).  Due to the
+coherently address ground states and single orbital excitations of up
+to five atoms per lattice site \cite{goban2018emergence}.  Due to the
 SU($N$) symmetry of inter-atomic interactions, manifest in the fact
-that all coupling constants are independent of nuclear spin, we can
-write all effective $M$-body Hamiltonians which address these states
-in the form
+that all coupling constants are independent of nuclear spin, a
+restriction of the $M$-body Hamiltonians $H_M=\sum_p H_M^{(p)}$ to the
+subspace of experimentally addressed states takes the form
 \begin{align}
   H_M = \sum_{\abs{\set{\mu_j}}=M}
-  \p{U_{M,\g} \prod_{\alpha=1}^M \n_{\mu_\alpha,\g}
-    + U_{M,+} \n_{\mu_1,\e} \prod_{\alpha=2}^M \n_{\mu_\alpha,\g}
+  \p{U_{M,\g} \n_{\mu_1,\g} \n_{\mu_2,\g}
+    + U_{M,+} \n_{\mu_1,\e} \n_{\mu_2,\g}
     + U_{M,-} \c_{\mu_1,\g}^\dag \c_{\mu_2,\e}^\dag
-    \c_{\mu_2,\g} \c_{\mu_1,\e} \prod_{\alpha=3}^M \n_{\mu_\alpha,\g}},
+    \c_{\mu_2,\g} \c_{\mu_1,\e}}
+  \prod_{\alpha=3}^M \n_{\mu_\alpha,\g},
   \label{eq:H_M}
@@
 near-term quantum simulations of SU($N$) spin models and lattice field
-theories.  Working in the deep-lattice limit, we have derived a
-low-energy effective theory of such atoms.  This theory exhibits
-emergent multi-body interactions that inherit the SU($N$) symmetry of
-the atoms' bare, two-body interactions.  When considering the
-low-lying orbital excitations of our effective theory, the SU($N$)
+theories.  Working in the deep-lattice limit and the experimental
+regime of at most one atom occupying each nuclear spin state on any
+lattice site, we have derived a low-energy effective theory for such
+atoms.  Our theory exhibits emergent multi-body interactions that
+inherit the SU($N$) symmetry of atoms' bare, two-body interactions.
+Considering a restriction of our theory to the subspace of at most one
+orbital excitation per lattice site, we found that the SU($N$)
 symmetry of $M$-body Hamiltonians allowed us to express them in a
@@
 \appendix
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Summary of experimental procedures}
-\label{sec:experiment_summary}
-
-The work in this paper is closely tied to the experimental work in
-ref.~\cite{goban2018emergence}; for quick reference, we here provide a
-short summary of the experimental procedures used therein.  The
-experiment begins by preparing a degenerate gas of $10^4$-$10^5$
-(fermionic) ${}^{87}$Sr atoms, in a uniform mixture of their ten
-nuclear spin states, at $\sim0.1$ of their fermi temperature (10 to 20
-nanokelvin) \cite{campbell2017fermidegenerate, marti2018imaging}.
-This gas is loaded into a primitive cubic optical lattice at the
-``magic wavelength'' for which both ground (${}^1S_0$) and
-first-excited (${}^3P_0$) electronic orbital states of the atoms
-experience the same lattice potential.  Lattice depths along the
-principal axes of the lattice are comparable in magnitude, with a
-geometric mean that can be varied from 30 to 80 $E_{\t{R}}$, where
-$E_{\t{R}}\approx3.5\times2\pi~\t{kHz}$ is the lattice photon recoil
-energy of the atoms.  These lattice depths are sufficiently large as
-to neglect tunneling on the time scales relevant to the experiment.
-Atoms are then addressed by an external (``clock'') interrogation
-laser with an ultranarrow (26 mHz) linewidth, detuned by $\Delta$ from
-the single-atom ${}^1S_0-{}^3P_0$ transition frequency.  When
-$\Delta=0$, non-interacting atoms in singly-occupied lattice sites
-thus undergo Rabi oscillations.  After a chosen time, the experiment
-removes all ground-state (${}^1S_0$) atoms from the lattice and uses
-absorption imaging to identify the remaining excited-state (${}^3P_0$)
-atoms.  As discussed in section \ref{sec:spectra}, a similar procedure
-with $\Delta\ne0$ leads to the spectrum shown in figure
-\ref{fig:sweep}, thereby constituting a measurement of the many-body
-atomic excitation spectrum.  The many-body atomic excitation spectrum
-is different from the single-body atomic excitation spectrum due to
-the presence of inter-atomic interactions.  We note that this
-spectroscopic protocol addresses only singly-excited orbital states of
-any lattice site.  Doubly-excited states are manifestly off-resonant
-for single-photon absorption processes, while multi-photon absorption
-processes are always off-resonant due to
-\begin{enumerate*}
-\item the non-linearity (on the scale of kHz) of on-site electronic
-  orbital excitation energies (resulting from non-uniform coupling
-  constants $G_X$), combined with
-\item the ultranarrow linewidth (on the scale of mHz) of the
-  interrogation laser.
-\end{enumerate*}
 
@@
 \end{align}
-where $H_{\t{eff}}^{(p)}$ is order $p$ in $V$.  Letting
-$\Q_0\equiv1-\P_0$ denote the projector onto $\E_0$ and $X$ denote any
-operator on $\H$, we define the superoperators
+where $H_{\t{eff}}^{(p)}$ is order $p$ in $V$.  Letting $\P_0$ denote
+the projector onto $\G_0$, $\Q_0\equiv\1-\P_0$ denote the projector
+onto $\E_0$, and $X$ denote any operator on $\H$, we define the
+superoperators
 \begin{align}
@@
 fermions in an optical lattice were performed in
-refs.~\cite{buchler2010microscopic} and \cite{wall2013strongly} using
+refs.~\cite{wall2013strongly} and \cite{buchler2010microscopic} using
 a two-channel model of a Feshbach resonance, yielding prescriptions
@@
   &&
-  K_{\t{HO}} \equiv \int d^3x~ \abs{\phi_0^{\t{HO}}}^4,
+  K_{\t{HO}} \equiv \int \d^3x~ \abs{\phi_0^{\t{HO}}}^4,
   \label{eq:harmonic_gamma}
@@
 scattering length $a_{\t{eff}}$ is given in terms of the zero-momentum
-scattering length $a$ by \cite{blume2002fermi, chin2010feshbach}
+scattering length $a$ by \cite{giorgini2008theory, blume2002fermi,
+  flambaum1999analytical}
 \begin{align}
@@
   H_{\t{int}}'
-  \equiv \f12 \sum {G'}^{qr}_{st} \int d^3 x~d^3y~\delta\p{z}
+  \equiv \f12 \sum {G'}^{qr}_{st} \int \d^3 x~d^3y~\delta\p{z}
   \sp{\hat\psi_{\mu q}^\dag\p{x} \hat\psi_{\nu r}^\dag\p{y}}
@@
 \begin{align}
-  K' \equiv \f12 \int d^3x~
+  K' \equiv \f12 \int \d^3x~
   \Re\sp{\p{\phi_0^*}^2

### 2019-02-07 multi_body_interactions/papers/theory/reviews_NJP/revisions.tex
added file summarizing all revisions to the NJP paper

\documentclass[preprint]{revtex4-1}

% linking references
\usepackage{hyperref}
\hypersetup{
  breaklinks=true,
  colorlinks=true,
  linkcolor=blue,
  urlcolor=cyan,
}

% general physics / math packages and commands
\usepackage{physics,amsmath,amssymb,braket,dsfont,bm}
\renewcommand{\t}{\text} % text in math mode
\newcommand{\f}{\dfrac} % shorthand for fractions
\newcommand{\p}[1]{\left(#1\right)} % parenthesis
\renewcommand{\sp}[1]{\left[#1\right]} % square parenthesis
\renewcommand{\set}[1]{\left\{#1\right\}} % curly parenthesis
\newcommand{\bk}{\braket} % shorthand for braket

\newcommand{\g}{\text{g}}
\newcommand{\e}{\text{e}}
\newcommand{\x}{\text{x}}
\newcommand{\y}{\text{y}}
\newcommand{\z}{\text{z}}

\renewcommand{\c}{\hat{c}}
\newcommand{\n}{\hat{n}}

\newcommand{\A}{\mathcal{A}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\G}{\mathcal{G}}
\renewcommand{\H}{\mathcal{H}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\K}{\mathcal{K}}
\renewcommand{\L}{\mathcal{L}}
\newcommand{\M}{\mathcal{M}}
\newcommand{\N}{\mathcal{N}}
\renewcommand{\O}{\mathcal{O}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\Q}{\mathcal{Q}}
\renewcommand{\S}{\mathcal{S}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\1}{\mathds{1}}

% inline lists
\usepackage[inline]{enumitem}
\setlist[enumerate,1]{label={(\roman*)}}

% color definitions (used in a figure)
\usepackage{xcolor} \newcommand{\blue}[1]{\textcolor{blue}{#1}}
\newcommand{\red}[1]{\textcolor{red}{#1}}
\newcommand{\green}[1]{\textcolor{green}{#1}}
\definecolor{lightblue}{RGB}{31,119,180}
\definecolor{orange}{RGB}{255,127,14}
\definecolor{green}{RGB}{44,160,44}
\definecolor{lightred}{RGB}{214,39,40}

% for strikeout text
% normalem included to prevent underlining titles in the bibliography
\usepackage[normalem]{ulem}

\begin{document}

Dear editors,

In this document, we summarize and comment on all revisions made to
our manuscript since its initial submission.  Revisions made in
response to comments and recommendations made by Referee 1 (Referee 2)
are numbered R1.$X$ (R2.$X$).  In addition, we summarize a few minor
revisions made to refine wording or fix errors in the original
manuscript; these revisions are numbered R3.$X$.

Excerpts from the referee reports are written in \blue{blue}, excerpts
from our original manuscript are written in \red{red}, and new text in
the revised version of our manuscript is written in \green{green}.
All page, line, and reference numbers generally refer to those in our
original manuscript.

\section*{Revisions prompted by Referee 1}

\begin{enumerate}[label=(R1.\arabic*)]
\item
\end{enumerate}

\section*{Revisions prompted by Referee 2}

\begin{enumerate}[label=(R2.\arabic*)]
\item
\end{enumerate}

\section*{Miscellaneous revisions}

\begin{enumerate}[label=(R3.\arabic*)]
\item Page 3, line 24 of our original manuscript previously said:

  \red{Last year (2017), a new generation of OLCs became operational
    at JILA}

  Reflecting the fact that it is now 2019, the new version of the
  manuscript says:

  \green{More recently (2017), a new generation of OLCs became
    operational at JILA}

\item Page 5, line 40 of our original manuscript previously said:

  \red{motional state $n\in\mathbb{N}_0^3$ with nuclear spin
    $\mu\in\set{-I,I+1,\cdots,I}$}

  This text contains a sign error; in the version of our manuscript,
  this text reads:

  \green{motional state $n\in\mathbb{N}_0^3$ with nuclear spin
    $\mu\in\set{-I,-I+1,\cdots,I}$}

\item Page 5, lines 43-44 of our original manuscript previously said:

  \red{In a harmonic trap approximation we would have
    $E_n=\p{3/2+n_x+n_y+n_z}\omega$}

  For clarity, we have modified this text so that all subscripts
  denoting free variables (e.g.~the $n$ in $E_n$, as $n$ can here be
  any element of $\mathbb{N}_0^3$) are italic, while all subscripts
  denoting fixed symbols (e.g.~the $x$ in $n_x$, as here $x$ denotes a
  fixed spatial axis) are in roman font.  The new text thus reads:

  \green{In a harmonic trap approximation we would have
    $E_n=\p{3/2+n_\x+n_\y+n_\z}\omega$}

  We have similarly replaced all appearances of $m_A$ (denoting the
  mass of a single atom) with $m_{\text{A}}$; all $E_R$ (denoting the
  lattice photon recoil energy) with $E_{\t{R}}$; all
  $k_L,k_L^x,k_L^y,k_L^z$ (denoting the wavenumber of the lattice
  laser and its projection onto three principal axes) by
  $k_{\text{L}}, k_{\text{L}}^\x, k_{\text{L}}^\y, k_{\text{L}}^\z$;
  all $\U_x,\U_y,\U_z$ (denoting lattice depths along principal axes)
  respectively by $\U_\x,\U_\y,\U_\z$; and all
  $T_\mu^z,T_\mu^x,S_{\N X}^z,S_{\N X}^x$ (denoting single-atom and
  many-body pseudospin operators) respectively by
  $T_\mu^\z,T_\mu^\x,S_{\N X}^\z,S_{\N X}^\x$.

  Finally, we have fixed a few discrepancies in our general convention
  that multiple subscripts on a given mathematical symbol should only
  be separated by a comma if at least one of these subscripts has a
  fixed value (as e.g.~in $K^{k\ell}_{mn}$, $K^{\ell,3}_{2,5}$,
  $\c_{\mu n}$, and $\c_{\mu,\g}$).  To this end, in our last appendix
  (page 40, line 17) we have replaced all instances of $U_{M,X}$ and
  $U_{M,X}^{(p)}$ respectively with $U_{MX}$ and $U_{MX}^{(p)}$.

\item On page 7, line 17 we previously had the text:

  \red{The existence of such a gauge is guaranteed by the analytic
    properties of Wannier orbitals $\phi_{in}$ [42].}

  With the exception of this sentence, our manuscript has generally
  reserved the word ``orbital'' to refer to electronic degrees of
  freedom of atoms, and ``motional state'' to refer to atoms' spatial
  degrees of freedom.  To avoid potential confusion, we have therefore
  modified the above sentence to read:

  \green{The existence of such a gauge is guaranteed by the analytic
    properties of Wannier wavefunctions $\phi_{in}$ [42].}

\item The last two lines of page 10 and line 26 of page 11 previously
  said:

  \red{Effective $M$-body interaction terms at order $p$ in
    $H_{\t{int}}$ can be represented by directed, weakly connected
    graphs (i.e.~which are connected when ignoring edge direction)
    containing $p$ internal vertices.}

  Though the final expressions in our perturbation theory can indeed
  be represented by weakly connected graphs, intermediate terms that
  cancel out may be represented by graphs that are not connected, as
  e.g.~in Eq.~(35) on page 17.  For clarity, we have therefore
  modified the above text to read:

  \green{Effective $M$-body interaction terms at order $p$ in
    $H_{\t{int}}$ can be represented by directed graphs containing $p$
    internal vertices.}

\item To clean up repetitive wording, we have replaced the sentence
  (page 12, line 45):

  \red{To leading order in the coupling constants, it follows that the
    counter-terms are second order in the coupling constants,
    i.e. $\tilde G\sim G^2$.}

  with:

  \green{To leading order, it follows that the counter-terms are
    second order in the coupling constants, i.e.~$\tilde G\sim G^2$.}

\item To make our first appendix more self-contained, we have defined
  $\P_0$ therein, such that the text (page 30, line 38):

  \red{Letting $\Q_0\equiv1-\P_0$ denote the projector onto $\E_0$ and
    $X$ ...}

  now reads:

  \green{Letting $\P_0$ denote the projector onto $\G_0$,
    $\Q_0\equiv\1-\P_0$ denote the projector onto $\E_0$, and $X$ ...}

\item Reference 34 of the original manuscript has been published since
  the original submission of our manuscript.  We have updated this
  reference accordingly in the bibliography of our revised manuscript.

\end{enumerate}

\end{document}

### 2019-02-08 multi_body_interactions/papers/theory/reviews_NJP/referee_1_response.tex
finished first draft of response to referee 1 from NJP

@@
 
+% for feynman diagrams
+\usepackage{tikz,tikz-feynman}
+\tikzset{
+  baseline = (current bounding box.center)
+}
+\tikzfeynmanset{
+  compat = 1.1.0,
+  every feynman = {/tikzfeynman/small}
+}
+\newcommand{\shrink}[1]{\scalebox{0.8}{#1}} % for smaller diagrams
+
 % color definitions (used in a figure)
@@
 All page, line, and reference numbers generally refer to those in our
-original manuscript.
+original manuscript, unless explicitly stated otherwise.
 
@@
   more accessible, and that we have addressed shortcomings identified
-  by the reviewer to a satisfactory degree.
+  by the referee to a satisfactory degree.
 
@@
 
-  \label{pt:limitations}
-
   We thank the referee for pointing out the potential to mislead
@@
 
-  In the introduction, we have modified the following sentence (in the
-  first paragraph discussing our technical contributions) (page 4,
-  line 22):
+  We then added explicit mention of ground-state preparation to the
+  introduction, replacing (page 4, line 22):
 
@@
 
-  which now reads:
+  by:
 
@@
 
-  We have also rearranged and rewritten some of the following
-  paragraph (page 4, line 33):
+  In order to motivate a later restriction to one orbital excitation
+  per lattice site, we have rearranged and rewritten some of the next
+  paragraph in the introduction, changing the text (page 4, line 33):
 
@@
   between Eqs.~(8) and (9), as discussed later in point
-  \ref{pt:confusion}, and have replaced the following text (page 7,
-  line 48):
+  \ref{pt:confusion}, and have made our technical contributions more
+  explicit in the preview of our main results (in section II),
+  replacing (page 7, line 48):
 
@@
     Hamiltonian can be written in the form
-    \begin{align}
+    \begin{align*}
       H_{\t{int}}^{\t{eff}} = \sum_{M=2}^{2I+1} H_M, \tag{9}
-    \end{align}
+    \end{align*}
     where $H_M$ is an $M$-body Hamiltonian, $I$ is the total nuclear
@@
 
-  with
+  by:
 
@@
     Hamiltonian can be written in the form
-    \begin{align}
+    \begin{align*}
       H_{\t{int}}^{\t{eff}} = \sum_{M=2}^{2I+1} \sum_{p\ge1} H_M^{(p)},
       \tag{9}
-    \end{align}
+    \end{align*}
     where $H_M^{(p)}$ is an $M$-body Hamiltonian of order $p$ in the
@@
 
-  In section IV (page 18), we have changed the first sentence from
-  (page 18, line line 45):
+  In section IV (page 18), we have similarly changed the first
+  sentence from (page 18, line line 45):
 
@@
 
-  Finally, we have modified the following text in our summary (page
-  28, line 40):
+  Finally, we have modified some text in our summary to emphasize that
+  our main theory is general with respect to orbital excitations, and
+  that the restriction to one orbital excitation per lattice site is
+  made to facilitate comparison with experiments; specifically, we
+  have changed (page 28, line 40):
 
@@
 
-  To reflect the addition of new text, we have modified the
+  To reflect this change to section II, we have modified the
   description of section II on page 4, line 52, previously:
@@
   \green{In section II we summarize the experimental procedures
-    relevant to our work, provide an overview one- and two-body
-    physics of ultracold atoms on a lattice, and preview our main
+    relevant to our work, provide an overview of the one- and two-body
+    physics of ultracold atoms in a deep lattice, and preview our main
     technical results.}
@@
 
-  We have re-written this paragraph in its entirety, replacing it with
-  the following text:
+  We have re-written this paragraph in its entirety to walk through
+  the arguments and explanations therein more carefully, replacing it
+  with the following text:
 
@@
   Accordingly, we have modified language in section III that makes
-  reference to the re-written paragraph, changing (page 9, line 45):
+  reference to the re-written paragraph above, changing (page 9, line
+  45):
 
@@
 
-  We hope that the new text is more readable and easier to understand.
+  We hope that the new text between Eqs.~(8) and (9) is more readable
+  and easier to understand.
 
@@
 
-  \red{\begin{align} H_M = \sum_{\abs{\set{\mu_j}}=M} \p{U_{M,\g}
+  \red{\begin{align*} H_M = \sum_{\abs{\set{\mu_j}}=M} \p{U_{M,\g}
         \prod_{\alpha=1}^M \n_{\mu_\alpha,\g} + U_{M,+} \n_{\mu_1,\e}
@@
       \tag{10}
-    \end{align}
+    \end{align*}
     where $\c_{\mu s}\equiv\c_{0,\mu s}$ is a fermionic annihilation
@@
 
-  \green{\begin{align}
+  \green{\begin{align*}
       H_M = \sum_{\abs{\set{\mu_j}}=M}
@@
       \tag{10}
-    \end{align}
+    \end{align*}
     where $H_2^{(\mu_1,\mu_2)}$ is a two-body Hamiltonian addressing
@@
 
-  and have modified Eq.~(42) to take the explicit form advertised in
-  the new Eq.~(10), changing (page 18, line 54):
-
-  \red{\begin{align}
+  and we have modified Eq.~(42) to take the explicit form advertised
+  in the new version of Eq.~(10), changing (page 18, line 54):
+
+  \red{\begin{align*}
       H_M = \sum_{\abs{\set{\mu_j}}=M}
@@
       \tag{42}
-    \end{align}}
+    \end{align*}}
 
@@
       \prod_{\alpha=3}^M \n_{\mu_\alpha,\g},
-      \tag{42}
-    \end{align}}
+      \tag{44}
+    \end{align*}}
+
+  Note that the new equation number is 44 due to the addition of two
+  new equations in response to point \ref{pt:divergence} below.
 
@@
 
+  \label{pt:pseudo-potential}
+
   We thank the referee for their attention to detail, and for pointing
@@
   $s$-wave interactions by our expression of Eq.~(2).  We have
-  therefore added the following paragraph to section II of our
-  manuscript, following the re-statement of Eq.~(2) in Eq.~(5)
-  (i.e.~the following text has been inserted in between lines 35 and
-  36 of page 6):
+  therefore added the a paragraph to section II, following the
+  re-statement of Eq.~(2) in Eq.~(5).  Specifically, the following
+  text has been inserted in between lines 35 and 36 of page 6:
 
@@
     it convenient to renormalize the coupling constants in the
-    effective theory developed in section III.  Second, (5) is only
-    the first term in a low-energy expansion of two-body interactions
-    in effective field theory, which generally includes additional
-    terms containing derivatives of field operators.  Derivative terms
-    correspond to the dependence of two-body scattering on the
-    relative momentum $k$ of particles involved, with $k\to0$ in the
-    zero-energy limit.  In the present case of $s$-wave scattering,
-    the leading dependence of the two-body interaction Hamiltonian on
-    the relative momentum $k$ can be captured by use of an
-    energy-dependent pseudo-potential, which amounts to using a
-    $k$-dependent effective scattering length [46].  This effective
-    scattering length can be determined by expanding the $s$-wave
-    collisional phase shift in powers of the relative momentum $k$
-    [45, 47].  Details of this expansion will depend on the
-    characteristic length scale of finite-range interactions.  In our
-    work, these corrections to (5) will be relevant only for the
+    effective theory developed in section III.  The choice of method
+    to regulate divergences has no effect on the underlying physics.}
+
+  \green{Second, (5) is only the first term in a low-energy expansion
+    of two-body interactions in effective field theory, which
+    generally includes additional terms containing derivatives of
+    field operators.  Derivative terms correspond to the dependence of
+    two-body scattering on the relative momentum $k$ of particles
+    involved, with $k\to0$ in the zero-energy limit.  In the present
+    case of $s$-wave scattering, the leading dependence of the
+    two-body interaction Hamiltonian on the relative momentum $k$ can
+    be captured by use of an energy-dependent pseudo-potential, which
+    amounts to using a $k$-dependent effective scattering length [46].
+    This effective scattering length can be determined by expanding
+    the $s$-wave collisional phase shift in powers of the relative
+    momentum $k$ [45, 47].  Details of this expansion will depend on
+    the characteristic length scale of finite-range interactions.  In
+    our work, these corrections to (5) will be relevant only for the
     calculation of two-body interaction energies, appearing at third
@@
     essentially the same as that used for similar calculations in
-    refs.~[35--37].}
-
-  Expanding on our use of renormalization, we have changed the
-  following text in section III B (page 12, line 15):
+    refs.~[35--37].  While this approach does not provide insight into
+    the microscopic nature of atomic interactions, it is suitable for
+    the phenomenological description of these interactions, and
+    particular for our eventual development of a low-energy effective
+    theory.}
+
+  Expanding on our use and method of renormalization, we have changed
+  the following text in section III B (page 12, line 15):
 
@@
     divergences ultimately appear due to our use of effective field
-    theory to describe inter-atomic interactions in (2) and (5),
+    theory to describe inter-atomic interactions in (2), (5), and (7),
     rather than a detailed microscopic description of two-atom
@@
     constants by introducing counter-terms $\tilde G_X$ into the
-    effective interaction Hamiltonians via $G_X\to G_X+\tilde G_X$.
-    The values of counter-terms can be fixed by enforcing that $G_X$
-    is equal to the net two-body interaction strength, which has the
-    added benefit of allowing us to express the effective two-body
-    interactions in terms of net, physically observable two-body
-    coupling constants on the right-hand side of (15), rather than in
-    terms of long sums at all orders of the bare coupling constants.}
+    interaction Hamiltonian.}
+
+  We further discuss our method of renormalization later in point
+  \ref{pt:renormalization}.
 
@@
 
-  \green{Second, our result in (17) neglects the effect of
+  \green{Finally, our result in (17) neglects the effect of
     momentum-dependent two-body scattering.  When the effective range
@@
 
-\item Concerning the effective theory corrections to many-body energy
-  eigenstates:
+\item Concerning corrections to many-body energy eigenstates in our
+  effective theory:
 
@@
   Hamiltonian $H_{\t{eff}}$ constructed from $H$ with a unitary
-  transformation, there are many different sets of energies and
-  eigenstates to keep track of.
-
-  To help clarify the relationship between $H_0$, $H$, and
-  $H_{\t{eff}}$, as well as their respective eigenstates and spectra,
-  we have re-written our opening to section III, replacing the
-  following paragraphs (page 8, line 36):
+  transformation, there are different sets of energies and eigenstates
+  to keep track of.
+
+  To help clarify the relationship between these different
+  Hamiltonians, their eigenstates, and spectra, we have re-written our
+  opening to section III, replacing the first two paragraphs (page 8,
+  line 36):
 
@@
     AEAs on a lattice is not diagonal with respect to single-particle
-    motional state indices.  The problem of determining interacting
-    atoms' orbital excitation spectrum therefore nominally involves
-    all atomic motional degrees of freedom.  At zero temperature,
-    however, each orbital state of interacting atoms is associated
-    with a single motional ground state.  In order to compute an
-    orbital excitation spectrum at zero temperature, in principle we
-    need to identify these motional ground states.  We can then ignore
-    all excited motional states, which will be neither thermally
-    occupied nor externally interrogated.  Such a procedure would
-    drastically reduce the dimensionality of the Hilbert space
-    necessary to describe the atoms, thereby greatly simplifying any
-    description of the atoms' orbital spectrum and dynamics.  In
-    practice, however, identifying the motional ground states of
-    interacting atoms and writing down a Hamiltonian restricted to
-    this subspace is a very difficult process to carry out
-    analytically.}
+    motional state indices (e.g.~$n\in\mathbb{N}_0^3$).  The problem
+    of determining interacting atoms' orbital excitation spectrum
+    therefore nominally involves all atomic motional degrees of
+    freedom.  At zero temperature, however, each orbital state of a
+    collection of interacting atoms is associated with a single
+    motional ground state.  In order to compute an orbital excitation
+    spectrum at zero temperature, in principle we need to identify
+    these motional ground states.  We can then ignore all excited
+    motional states, which will be neither thermally occupied nor
+    externally interrogated.  Such a procedure would drastically
+    reduce the dimensionality of the Hilbert space necessary to
+    describe the atoms, thereby greatly simplifying any description of
+    the atoms' orbital spectrum and internal (i.e.~nuclear and
+    orbital) dynamics.  In practice, however, identifying the motional
+    ground states of interacting atoms and writing down a Hamiltonian
+    restricted to this subspace is a very difficult process to carry
+    out analytically.}
 
@@
 
-  We have also modified the text pointed out by the referee, changing
-  (page 10, line 19):
+  We have also modified the particular text pointed out by the
+  referee, changing (page 10, line 19):
 
@@
   \green{Finally, the effective theory involves no corrections to the
-    non-interacting many-body energy eigenstates; the purpose of the
-    effective Hamiltonian $H_{\t{eff}} = H_0 + H_{\t{int}}^{\t{eff}}$
-    is to reproduce, on the eigenstates of the non-interacting
-    Hamiltonian $H_0$ within $\H_{\t{ground}}^{\t{single}}$, the
-    spectrum of the interacting Hamiltonian $H = H_0 + H_{\t{int}}$ on
-    $\H_{\t{ground}}^{\t{multi}}$.  Modifying the eigenstates of the
-    non-interacting Hamiltonian $H_0$ on
+    non-interacting many-body energy eigenstates; the purpose of
+    constructing the effective Hamiltonian
+    $H_{\t{eff}} = H_0 + H_{\t{int}}^{\t{eff}}$ is to reproduce, on
+    the known eigenstates of the non-interacting Hamiltonian $H_0$
+    within $\H_{\t{ground}}^{\t{single}}$, the spectrum of the
+    interacting Hamiltonian $H = H_0 + H_{\t{int}}$ on
+    $\H_{\t{ground}}^{\t{multi}}$.  ``Correcting'' the eigenstates of
+    the non-interacting Hamiltonian $H_0$ on
     $\H_{\t{ground}}^{\t{single}}$ thus invalidates the effective
@@
 
+  We agree with the referee that a microscopic derivation of our
+  effective theory based on a regularized pseudo-potential would
+  provide additional insights, but respectfully disagree that doing so
+  is worthwhile for our manuscript.  Microscopic treatment of
+  interactions is generally a difficult task; we do not know of any
+  analytical descriptions of even {\it two}-body interactions of atoms
+  on a lattice (as exist e.g.~in the case of a harmonic trap,
+  considered in
+  ref.~[\href{https://link.springer.com/article/10.1023\%2FA\%3A1018705520999}{53}]),
+  nevermind multi-body interactions that appear in an effective theory
+  of the two-body case.  Effective field theory and renormalization,
+  meanwhile, are standard and powerful tools for studying interacting
+  quantum systems.  These tools are therefore well-suited to our goal
+  at hand: to interpret the results of experiments exhibiting
+  signatures of multi-body interactions, and build towards the use of
+  such experiments to perform interesting quantum simulations.  We
+  discussed our use of effective field theory and renormalization in
+  our response to point \ref{pt:pseudo-potential}, commenting on our
+  particular choice of methods in a pair new sentences inserted
+  between Eqs.~(5) and (6):
+
+  \green{We note that our approach of using an unregularized contact
+    potential, renormalizing coupling constants, and separately
+    accounting for momentum-dependent scattering is essentially the
+    same as that used for similar calculations in refs.~[35--37].
+    While this approach does not provide insight into the microscopic
+    structure of inter-atomic interactions, it is suitable for the
+    phenomenological description of these interactions, and particular
+    for our eventual development of a low-energy effective theory.}
+
+  We hope that the referee understands our position on this matter,
+  and does not consider a microscopic treatment of interactions to be
+  compulsory for the publication of our manuscript.
+
+\item Concerning the diagram in Eq.~(22):
+
+  \blue{The referee tried to follow the authors approach, but was
+    stuck with the following two problems:}
+
+  \blue{(1) It is unclear to the referee, why the diagram in Eq.(22)
+    converges. For large n and m motional state reduces to free
+    particles with energy $\sim$ n\^{}2 and m\^{}2 as the energy is
+    much higher than the lattice. In addition, the overlap K become a
+    constant. Then, a dimensional analysis shows that this term
+    diverges logarithmically. It is a significant short coming of the
+    manuscript that the authors fail to provide a detailed discussion
+    of potential diverging terms. Therefore, it is unclear to the
+    referee, whether the chosen approach is well defined.}
+
+  \label{pt:divergence}
+
+  We thank the referee for their careful consideration of our theory,
+  for pointing out a potential concern about divergences.  The
+  argument for a divergence in our theory essentially concerns a
+  quantity in Eq.~(30) on page 16:
+  \begin{align*}
+    \alpha_3^{(3)} \equiv \sum_{\substack{\ell+m>0\\\ell+n>0}}
+    \f{K_{\ell m} K^m_n K_{\ell n}}{\p{E_\ell+E_m}\p{E_\ell+E_n}},
+  \end{align*}
+  for $\ell,m,n\in\mathbb{N}_0^3$ with
+  e.g.~$\ell\equiv\p{\ell_\x,\ell_\y,\ell_\z}$ and $E_\ell$ equal to
+  the motional energy of a non-interacting atom in motional state
+  $\ell$.  The referee correctly recognizes that when $\ell$ is large,
+  it indexes the state of a free particle with an asymptotic energy
+  $E_\ell\sim\ell^2\equiv\ell_\x^2+\ell_\y^2+\ell_\z^2$.  If spatial
+  overlap integrals (e.g.~$K_{\ell m}$) become constant for large
+  $\ell,m,n$, the sum above asymptotically takes the form
+  \begin{align*}
+    \alpha_3^{(3)} \stackrel{?}{\sim}
+    \int \f{\d^3\ell~\d^3m~\d^3n}{\p{\ell^2+m^2}\p{\ell^2+n^2}},
+  \end{align*}
+  which diverges by dimensional analysis.
+
+  The problem with the above argument is that the overlap integral
+  $K_{\ell m}$ does not become constant at large $\ell,m$, but rather
+  becomes sharply peaked at $\ell\approx m$.  We address and explain
+  this point in the following text, which we have appended to section
+  III D after Eq.~(32) on page 16:
+
+  \green{Note that although (22) is a loop diagram, the associated
+    factors $\alpha_3^{(3)}$ and $\alpha_5^{(3)}$ in $H_3^{(3)}$ are
+    finite.  At large $n$, atoms become free particles for which $n$
+    essentially indexes discrete momentum states.  These atoms thus
+    have an energy which asymptotically scales as
+    $E_n\sim n^2\equiv n_\x^2+n_\y^2+n_\z^2$.  Furthermore, the
+    oscillatory behavior of atomic wavefunctions with increasing
+    motional state indices $\ell,m$ implies that the overlap integral
+    $K_{\ell m}$ (and similarly $K^\ell_m$) becomes sharply peaked at
+    $\ell\approx m$ as $\ell$ and $m$ get large.  The asymptotic
+    behavior of $\alpha_3^{(3)}$ at large $\ell,m,n$ is therefore
+    \begin{align*}
+      \alpha_3^{(3)}
+      \sim \int \f{\d^3\ell~\d^3m~\d^3n}{\p{\ell^2+m^2}\p{\ell^2+n^2}}
+      ~\delta\p{\ell-m}\delta\p{\ell-n}
+      \sim \int \f{\d^3\ell}{\ell^4}
+      \sim \int_{\ell_{\t{min}}}^\infty \f{\d\ell}{\ell^2}
+      \sim \f1{\ell_{\t{min}}},
+      \tag{33}
+    \end{align*}
+    where in the last integral we changed to spherical coordinates,
+    and $\ell_{\t{min}}^2$ is the minimum value of $\ell^2$ for which
+    (i) the energy $E_\ell\sim\ell^2$, and (ii) the integral
+    expression in (33) is a good approximation to the corresponding
+    sum in (30).  Note that the introduction of $\ell_{\t{min}}$
+    amounts to neglecting a finite number of terms in the sum over
+    $\ell,m,n$ in (30), whose contribution to the value of
+    $\alpha_3^{(3)}$ is finite.  Convergence of $\alpha_5^{(3)}$ is
+    similarly guaranteed by the fact that the overlap integral $K_n$
+    does not asymptotically grow with increasing $n$, such that
+    $\alpha_5^{(3)}$ asymptotically behaves as
+    \begin{align*}
+      \alpha_5^{(3)} \sim \int \f{\d^3 n}{n^4}
+      \sim \int_{n_{\t{min}}}^\infty \f{\d n}{n^2}
+      \sim \f1{n_{\t{min}}},
+      \tag{34}
+    \end{align*}
+    where $n_{\t{min}}$ is defined similarly to $\ell_{\t{min}}$.}
+
+  We hope that this new text addresses potential concerns about
+  divergences in a satisfactory manner.
+
+\item Concerning the last point:
+
+  \blue{(ii) The referee is a bit familiar with counter terms in
+    standard renormalization of field theories like QED. However,
+    there it is required that a counter term in each order
+    perturbation theory is introduced. In turn, the authors in the
+    present manuscript just introduce a single counter term. The
+    referee would expect that a careful theoretical manuscript
+    properly discusses the regularization and renormalization of the
+    theory as well as all diverging diagrams.}
+
+  \label{pt:renormalization}
+
+  We thank the referee for pointing out that our manuscript would
+  benefit from a careful explanation of our renormalization scheme.
+  Renormalization is a class of techniques, the details of which may
+  look different in the context of different field theories; worse
+  still, any chosen means to renormalize a field theory is generally
+  not unique.
+
+  In order to clarify our introduction and use of counter-terms, we
+  have replaced the following text (page 12, line 35):
+
+  \red{The values of these counter-terms are fixed by enforcing that
+    $G_X$ is equal to the net two-body interaction
+    strength. Diagrammatically, this renormalization condition is
+    \begin{align*}
+      \substack{
+        \shrink{
+          \begin{tikzpicture}
+            \begin{feynman}
+              \vertex (v);
+              \vertex[above left = of v] (f1);
+              \vertex[below left = of v] (f2);
+              \vertex[above right = of v] (f3);
+              \vertex[below right = of v] (f4);
+              \diagram* {
+                (f1) -- (v),
+                (f2) -- (v),
+                (v) -- (f3),
+                (v) -- (f4) };
+            \end{feynman}
+          \end{tikzpicture}}
+        \\ \O(G)}
+      + \substack{
+        \shrink{
+          \begin{tikzpicture}
+            \begin{feynman}
+              \vertex[crossed dot] (v) {};
+              \vertex[above left = of v] (f1);
+              \vertex[below left = of v] (f2);
+              \vertex[above right = of v] (f3);
+              \vertex[below right = of v] (f4);
+              \diagram* {
+                (f1) -- (v),
+                (f2) -- (v),
+                (v) -- (f3),
+                (v) -- (f4) };
+            \end{feynman}
+          \end{tikzpicture}}
+        \\ \O(\tilde G)}
+      - \substack{
+        \shrink{
+          \begin{tikzpicture}
+            \begin{feynman}
+              \vertex (v1);
+              \vertex[above left = of v1] (f1);
+              \vertex[below left = of v1] (f2);
+              \vertex[right = of v1] (v2);
+              \vertex[above right = of v2] (f3);
+              \vertex[below right = of v2] (f4);
+              \diagram* {
+                (f1) -- (v1),
+                (f2) -- (v1),
+                (v2) -- (f3),
+                (v2) -- (f4),
+                (v1) --[scalar, half left] (v2),
+                (v1) --[scalar, half right] (v2) };
+            \end{feynman}
+          \end{tikzpicture}}
+        \\ \O(G^2)}
+      + \cdots = \substack{
+        \shrink{
+          \begin{tikzpicture}
+            \begin{feynman}
+              \vertex (v);
+              \vertex[above left = of v] (f1);
+              \vertex[below left = of v] (f2);
+              \vertex[above right = of v] (f3);
+              \vertex[below right = of v] (f4);
+              \diagram* {
+                (f1) -- (v),
+                (f2) -- (v),
+                (v) -- (f3),
+                (v) -- (f4) };
+            \end{feynman}
+          \end{tikzpicture}}
+        \\ \O(G)}.
+      \tag{16}
+    \end{align*}
+    where we represent counter-terms by a crossed dot (i.e.~$\otimes$)
+    at a vertex. To leading order in the coupling constants, it
+    follows that the counter-terms are second order in the coupling
+    constants, i.e.~$\tilde G\sim G^2$. This finding implies that
+    counter-terms only come into play at next- to-leading order (NLO)
+    in the calculation of any effective $M$-body interaction
+    Hamiltonian.}
+
+  by a new paragraph with more detailed discussion of counter-terms
+  and the renormalization condition:
+
+  \green{The introduction of counter-terms is merely a formal
+    decomposition of the ``bare'' coupling constants $G_X^{\t{bare}}$
+    that are used in (15) as $G_X^{\t{bare}} = G_X + \tilde G_X$.  In
+    performing such a decomposition, we are free to choose the values
+    of $G_X$, which in turn fixes the values of
+    $\tilde G_X \equiv G_X^{\t{bare}} - G_X$.  For convenience, we can
+    choose the values of $G_X$ to be those of the physically observed
+    effective coupling constants on the right-hand side of (15).
+    Representing the new coupling constants $G_X$ by regular vertices
+    and the counter-terms $\tilde G_X$ by a crossed dot
+    (i.e.~$\otimes$), this choice leads to the {\it renormalization
+      condition}
+    \begin{align*}
+      \substack{
+        \shrink{
+          \begin{tikzpicture}
+            \begin{feynman}
+              \vertex (v);
+              \vertex[above left = of v] (f1);
+              \vertex[below left = of v] (f2);
+              \vertex[above right = of v] (f3);
+              \vertex[below right = of v] (f4);
+              \diagram* {
+                (f1) -- (v),
+                (f2) -- (v),
+                (v) -- (f3),
+                (v) -- (f4) };
+            \end{feynman}
+          \end{tikzpicture}}
+        \\ \O(G)}
+      + \substack{
+        \shrink{
+          \begin{tikzpicture}
+            \begin{feynman}
+              \vertex[crossed dot] (v) {};
+              \vertex[above left = of v] (f1);
+              \vertex[below left = of v] (f2);
+              \vertex[above right = of v] (f3);
+              \vertex[below right = of v] (f4);
+              \diagram* {
+                (f1) -- (v),
+                (f2) -- (v),
+                (v) -- (f3),
+                (v) -- (f4) };
+            \end{feynman}
+          \end{tikzpicture}}
+        \\ \O(\tilde G)}
+      - \substack{
+        \shrink{
+          \begin{tikzpicture}
+            \begin{feynman}
+              \vertex (v1);
+              \vertex[above left = of v1] (f1);
+              \vertex[below left = of v1] (f2);
+              \vertex[right = of v1] (v2);
+              \vertex[above right = of v2] (f3);
+              \vertex[below right = of v2] (f4);
+              \diagram* {
+                (f1) -- (v1),
+                (f2) -- (v1),
+                (v2) -- (f3),
+                (v2) -- (f4),
+                (v1) --[scalar, half left] (v2),
+                (v1) --[scalar, half right] (v2) };
+            \end{feynman}
+          \end{tikzpicture}}
+        \\ \O(G^2)}
+      + \cdots = \substack{
+        \shrink{
+          \begin{tikzpicture}
+            \begin{feynman}
+              \vertex (v);
+              \vertex[above left = of v] (f1);
+              \vertex[below left = of v] (f2);
+              \vertex[above right = of v] (f3);
+              \vertex[below right = of v] (f4);
+              \diagram* {
+                (f1) -- (v),
+                (f2) -- (v),
+                (v) -- (f3),
+                (v) -- (f4) };
+            \end{feynman}
+          \end{tikzpicture}}
+        \\ \O(G)}.
+      \tag{16}
+    \end{align*}
+    This renormalization condition has the benefit of allowing us to
+    express effective two-body interactions simply in terms of
+    physically observed effective two-body coupling constants, rather
+    than in terms of long sums at all order of the bare coupling
+    constants.}
+
+  We have also added a paragraph explaining where and why counter-term
+  diagrams should appear in our effective theory, in particular
+  explaining why counter-terms do not appear at leading order for
+  effective $M$-body interactions.  This new paragraph is inserted
+  between lines 23 and 24 on page 13:
+
+  \green{Second, the renormalization condition in (16) implies that
+    the counter terms $\tilde G_X$ are second order in the coupling
+    constants $G_X$, i.e.~$\tilde G_X\sim G_X^2$.  Although the
+    effective Hamiltonian expansions in (13) and (14) are organized in
+    powers of the interaction Hamiltonian $H_{\t{int}}$, the couplings
+    $G_X$ are the ``small'' parameters in which we can formally
+    organize a perturbation theory; more specifically, the formally
+    small quantities organizing our perturbation theory are two-body
+    ground-state interaction energies (proportional to the couplings
+    $G_X$) divided by the spectral gap of the single-atom Hamiltonian
+    $H_0$.  If $M$ atoms can only couple through terms represented by
+    a $p$-vertex diagrams for $p\ge p_M^{\t{min}}$, then the leading
+    order contribution to $M$-body interactions is order
+    $p_M^{\t{min}}$ in the couplings $G_X$.  If the same
+    $p_M^{\t{min}}$-vertex diagrams involve any counter-terms,
+    however, then these diagrams are at least order $p_M^{\t{min}}+1$
+    in the couplings $G_X$.  Counter-terms therefore only appear at
+    next-to-leading order (NLO) in the calculation of effective
+    $M$-body interactions.}
+
+  To reflect this added paragraph, we have modified the text on page
+  13, line 7:
+
+  \red{Before moving on to consider effective three-body interactions,
+    there are two comments we must make concerning the result in
+    (17).}
+
+  to read:
+
+  \green{Before moving on to consider effective three-body
+    interactions, there are a few comments we must make concerning
+    renormalization and the result in (17).}
+
+  We also addressed the subject of divergences and the reason for
+  their appearance in new text immediately following Eq.~(15) (page
+  12), as discussed earlier in point \ref{pt:pseudo-potential}.
+
+  We hope that these changes clarify any confusions about our
+  renormalization scheme, and in particular about the appearances (or
+  lack thereof) of counter-terms at different orders in our
+  perturbative effective theory.
+
+\item Finally:
+
+  \blue{In summary, the manuscript can not be published in the present
+    form without strongly improving the readability and in addition
+    provide a more detailed analysis of all diverging diagrams and a
+    proper discussion of their renormalization.}
+
+  We thank the referee for their extensive and detailed review of our
+  manuscript.  The referee brought up many good points, which we have
+  addressed by the addition of new discussions and the modification of
+  existing text.  In light of these revisions, we hope that the
+  referee deems our manuscript suitable for publication.  In
+  particular, we hope the referee considers our revised manuscript
+  more readable and accessible, with sufficient discussion and
+  clarification of the potential confusions and concerns brought up by
+  the referee.
+
 \end{enumerate}

### 2019-02-08 multi_body_interactions/papers/theory/paper.tex
finished first draft of response to referee 1 from NJP

@@
 \newcommand{\U}{\mathcal{U}}
+
 \newcommand{\1}{\mathds{1}}
@@
 \ref{sec:overview} we summarize the experimental procedures relevant
-to our work, provide an overview one- and two-body physics of
-ultracold atoms on a lattice, and preview our main technical results.
-In section \ref{sec:low_energy} we discuss our method for deriving a
-low-energy effective theory, provide a perturbative expansion for the
-net effective Hamiltonian, and compute all $M$-body Hamiltonians
-through third order in the low-energy effective theory.  We then
-analyze the low-lying excitations of the effective theory in section
-\ref{sec:spectra}, comparing spectral predictions with experimental
-measurements, and study the orbital-state dynamics of nuclear spin
-mixtures interrogated via Rabi spectroscopy.  Finally, we summarize
-and conclude our findings in section \ref{sec:summary}, and provide
-some discussion of future outlooks.
+to our work, provide an overview of the one- and two-body physics of
+ultracold atoms in a deep lattice, and preview our main technical
+results.  In section \ref{sec:low_energy} we discuss our method for
+deriving a low-energy effective theory, provide a perturbative
+expansion for the net effective Hamiltonian, and compute all $M$-body
+Hamiltonians through third order in the low-energy effective theory.
+We then analyze the low-lying excitations of the effective theory in
+section \ref{sec:spectra}, comparing spectral predictions with
+experimental measurements, and study the orbital-state dynamics of
+nuclear spin mixtures interrogated via Rabi spectroscopy.  Finally, we
+summarize and conclude our findings in section \ref{sec:summary}, and
+provide some discussion of future outlooks.
 
@@
 renormalize the coupling constants in the effective theory developed
-in section \ref{sec:low_energy}.  Second, \eqref{eq:H_int_fields} is
-only the first term in a low-energy expansion of two-body interactions
-in effective field theory, which generally includes additional terms
-containing derivatives of field operators.  Derivative terms
-correspond to the dependence of two-body scattering on the relative
-momentum $k$ of particles involved, with $k\to0$ in the zero-energy
-limit.  In the present case of $s$-wave scattering, the leading
-dependence of the two-body interaction Hamiltonian on the relative
-momentum $k$ can be captured by use of an energy-dependent
-pseudo-potential, which amounts to using a $k$-dependent effective
-scattering length \cite{blume2002fermi}.  This effective scattering
-length can be determined by expanding the $s$-wave collisional phase
-shift in powers of the relative momentum $k$ \cite{giorgini2008theory,
-  flambaum1999analytical}.  Details of this expansion will depend on
-the characteristic length scale of finite-range interactions.  In our
-work, these corrections to \eqref{eq:H_int_fields} will be relevant
-only for the calculation of two-body interaction energies, appearing
-at third order in the coupling constants $G_X$.  As we are primarily
-interested in $M$-body interactions for $M\ge3$, we defer this
-calculation to Appendix \ref{sec:momentum_dependence}.  We note that
-our approach of using an unregularized contact potential,
-renormalizing coupling constants, and separately accounting for
-momentum-dependent scattering is essentially the same as that used for
-similar calculations in refs.~\cite{johnson2012effective,
-  yin2014universal, johnson2009effective}.
+in section \ref{sec:low_energy}.  The choice of method to regulate
+divergences has no effect on the underlying physics.
+
+Second, \eqref{eq:H_int_fields} is only the first term in a low-energy
+expansion of two-body interactions in effective field theory, which
+generally includes additional terms containing derivatives of field
+operators.  Derivative terms correspond to the dependence of two-body
+scattering on the relative momentum $k$ of particles involved, with
+$k\to0$ in the zero-energy limit.  In the present case of $s$-wave
+scattering, the leading dependence of the two-body interaction
+Hamiltonian on the relative momentum $k$ can be captured by use of an
+energy-dependent pseudo-potential, which amounts to using a
+$k$-dependent effective scattering length \cite{blume2002fermi}.  This
+effective scattering length can be determined by expanding the
+$s$-wave collisional phase shift in powers of the relative momentum
+$k$ \cite{giorgini2008theory, flambaum1999analytical}.  Details of
+this expansion will depend on the characteristic length scale of
+finite-range interactions.  In our work, these corrections to
+\eqref{eq:H_int_fields} will be relevant only for the calculation of
+two-body interaction energies, appearing at third order in the
+coupling constants $G_X$.  As we are primarily interested in $M$-body
+interactions for $M\ge3$, we defer this calculation to Appendix
+\ref{sec:momentum_dependence}.  We note that our approach of using an
+unregularized contact potential, renormalizing coupling constants, and
+separately accounting for momentum-dependent scattering is essentially
+the same as that used for similar calculations in
+refs.~\cite{johnson2012effective, yin2014universal,
+  johnson2009effective}.  While this approach does not provide insight
+into the microscopic structure of inter-atomic interactions, it is
+suitable for the phenomenological description of these interactions,
+and particular for our eventual development of a low-energy effective
+theory.
 
@@
 
-\newpage
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 lattice is not diagonal with respect to single-particle motional state
-indices.  The problem of determining interacting atoms' orbital
-excitation spectrum therefore nominally involves all atomic motional
-degrees of freedom.  At zero temperature, however, each orbital state
-of interacting atoms is associated with a single motional ground
-state.  In order to compute an orbital excitation spectrum at zero
-temperature, in principle we need to identify these motional ground
-states.  We can then ignore all excited motional states, which will be
-neither thermally occupied nor externally interrogated.  Such a
-procedure would drastically reduce the dimensionality of the Hilbert
-space necessary to describe the atoms, thereby greatly simplifying any
-description of the atoms' orbital spectrum and dynamics.  In practice,
-however, identifying the motional ground states of interacting atoms
-and writing down a Hamiltonian restricted to this subspace is a very
-difficult process to carry out analytically.
+indices (e.g.~$n\in\mathbb{N}_0^3$).  The problem of determining
+interacting atoms' orbital excitation spectrum therefore nominally
+involves all atomic motional degrees of freedom.  At zero temperature,
+however, each orbital state of a collection of interacting atoms is
+associated with a single motional ground state.  In order to compute
+an orbital excitation spectrum at zero temperature, in principle we
+need to identify these motional ground states.  We can then ignore all
+excited motional states, which will be neither thermally occupied nor
+externally interrogated.  Such a procedure would drastically reduce
+the dimensionality of the Hilbert space necessary to describe the
+atoms, thereby greatly simplifying any description of the atoms'
+orbital spectrum and internal (i.e.~nuclear and orbital) dynamics.  In
+practice, however, identifying the motional ground states of
+interacting atoms and writing down a Hamiltonian restricted to this
+subspace is a very difficult process to carry out analytically.
 
@@
 corrections to the non-interacting many-body energy eigenstates; the
-purpose of the effective Hamiltonian
+purpose of constructing the effective Hamiltonian
 $H_{\t{eff}} = H_0 + H_{\t{int}}^{\t{eff}}$ is to reproduce, on the
-eigenstates of the non-interacting Hamiltonian $H_0$ within
+known eigenstates of the non-interacting Hamiltonian $H_0$ within
 $\H_{\t{ground}}^{\t{single}}$, the spectrum of the interacting
 Hamiltonian $H = H_0 + H_{\t{int}}$ on $\H_{\t{ground}}^{\t{multi}}$.
-Modifying the eigenstates of the non-interacting Hamiltonian $H_0$ on
-$\H_{\t{ground}}^{\t{single}}$ thus invalidates the effective theory.
+``Correcting'' the eigenstates of the non-interacting Hamiltonian
+$H_0$ on $\H_{\t{ground}}^{\t{single}}$ thus invalidates the effective
+theory.
 
@@
 similarly to the diagrams used to represent elements of the scattering
-matrix in field theory.  Effective $M$-body interaction terms at order
-$p$ in $H_{\t{int}}$ can be represented by directed graphs containing
-$p$ internal vertices.  Each internal vertex and its associated edges
-correspond respectively to a coupling constant and the associated
-field operators in $H_{\t{int}}$.  An example 2-vertex diagram
-representing an effective 3-body interaction term is provided in
-figure \ref{fig:diagram}.  All diagrams are read from left to right to
-construct a sequence of operators from right to left; the internal
-vertices of a diagram are thus strictly ordered, with the $n$-th
-internal vertex from the left corresponding to the $n$-th interaction
-Hamiltonian $H_{\t{int}}$ from the right in \eqref{eq:H_int_1_2} or
-\eqref{eq:H_int_3}.  Solid (dashed) lines represent field operators
-acting on the lowest (arbitrary) motional states.  Spatial overlap
-factors at each vertex are determined by the motional states of the
-edges which connect to (i.e.~field operators associated with) that
-vertex.  While it is possible to construct explicit rules for
-determining the energetic suppression factors (i.e.~from $\I$) of the
-term represented by a diagram, these factors are most easily
-determined by examination of the effective Hamiltonians in
-\eqref{eq:H_int_1_2} and \eqref{eq:H_int_3}.
+matrix in standard quantum field theory.  Effective $M$-body
+interaction terms at order $p$ in $H_{\t{int}}$ can be represented by
+directed graphs containing $p$ internal vertices.  Each internal
+vertex and its associated edges correspond respectively to a coupling
+constant and the associated field operators in $H_{\t{int}}$.  An
+example 2-vertex diagram representing an effective 3-body interaction
+term is provided in figure \ref{fig:diagram}.  All diagrams are read
+from left to right to construct a sequence of operators from right to
+left; the internal vertices of a diagram are thus strictly ordered,
+with the $n$-th internal vertex from the left corresponding to the
+$n$-th interaction Hamiltonian $H_{\t{int}}$ from the right in
+\eqref{eq:H_int_1_2} or \eqref{eq:H_int_3}.  Solid (dashed) lines
+represent field operators acting on the lowest (arbitrary) motional
+states.  Spatial overlap factors at each vertex are determined by the
+motional states of the edges which connect to (i.e.~field operators
+associated with) that vertex.  While it is possible to construct
+explicit rules for determining the energetic suppression factors
+(i.e.~from $\I$) of the term represented by a diagram, these factors
+are most easily determined by examination of the effective
+Hamiltonians in \eqref{eq:H_int_1_2} and \eqref{eq:H_int_3}.
 
@@
     understanding that the low-energy effective theory implicitly
-    addresses only single-atom motional ground states.}
-  \label{fig:diagram}
+    addresses only single-atom motional ground states.
+    \label{fig:diagram}}
 \end{figure}
@@
 appear due to our use of effective field theory to describe
-inter-atomic interactions in \eqref{eq:H_int_start} and
-\eqref{eq:H_int_fields}, rather than a detailed microscopic
-description of two-atom scattering.  Divergences of this sort are a
-generic feature of field theories, and can be dealt with using
-standard techniques such as renormalization.  We therefore renormalize
-our coupling constants by introducing counter-terms $\tilde G_X$ into
-the effective interaction Hamiltonians via $G_X\to G_X+\tilde G_X$.
-The values of counter-terms can be fixed by enforcing that $G_X$ is
-equal to the net two-body interaction strength, which has the added
-benefit of allowing us to express the effective two-body interactions
-in terms of net, physically observable two-body coupling constants on
-the right-hand side of \eqref{eq:H_2_expansion}, rather than in terms
-of long sums at all orders of the bare coupling constants.
-Diagrammatically, this renormalization condition is
+inter-atomic interactions in \eqref{eq:H_int_start},
+\eqref{eq:H_int_fields}, and \eqref{eq:H_int}, rather than a detailed
+microscopic description of two-atom scattering.  Divergences of this
+sort are a generic feature of field theories, and can be dealt with
+using standard techniques such as renormalization.  We therefore
+renormalize our coupling constants by introducing counter-terms
+$\tilde G_X$ into the interaction Hamiltonian.
+
+The introduction of counter-terms is merely a formal decomposition of
+the ``bare'' coupling constants $G_X^{\t{bare}}$ that are used in
+\eqref{eq:H_2_expansion} as $G_X^{\t{bare}} = G_X + \tilde G_X$.  In
+performing such a decomposition, we are free to choose the values of
+$G_X$, which in turn fixes the values of
+$\tilde G_X \equiv G_X^{\t{bare}} - G_X$.  For convenience, we can
+choose the values of $G_X$ to be those of the physically observed
+effective coupling constants on the right-hand side of
+\eqref{eq:H_2_expansion}.  Representing the new coupling constants
+$G_X$ by regular vertices and the counter-terms $\tilde G_X$ by a
+crossed dot (i.e.~$\otimes$), this choice leads to the {\it
+  renormalization condition}
 \begin{align}
@@
       \end{tikzpicture}}
-    \\ \O(G)},
+    \\ \O(G)}.
   \label{eq:counter_term}
 \end{align}
-where we represent counter-terms by a crossed dot (i.e.~$\otimes$) at
-a vertex. To leading order, it follows that the counter-terms are
-second order in the coupling constants, i.e.~$\tilde G\sim G^2$.  This
-finding implies that counter-terms only come into play at
-next-to-leading order (NLO) in the calculation of any effective
-$M$-body interaction Hamiltonian.  By construction, these
-counter-terms exactly cancel all terms beyond leading order in
-\eqref{eq:H_2_expansion}, which implies that the effective two-body
-interaction Hamiltonian is simply
+This renormalization condition has the benefit of allowing us to
+express effective two-body interactions simply in terms of physically
+observed effective two-body coupling constants, rather than in terms
+of long sums at all order of the bare coupling constants.  By
+construction, the counter-terms we have introduced exactly cancel all
+terms beyond leading order in \eqref{eq:H_2_expansion}, which implies
+that the effective two-body interaction Hamiltonian is simply
 \begin{align}
@@
 Before moving on to consider effective three-body interactions, there
-are two comments we must make concerning the result in \eqref{eq:H_2}.
-First, this effective two-body interaction takes the same form as the
-bare two-body interaction in \eqref{eq:H_int}, but without excited
+are a few comments we must make concerning renormalization and the
+result in \eqref{eq:H_2}.  First, this effective two-body interaction
+$H_2$ in \eqref{eq:H_2} takes the same form as the bare two-body
+interaction $H_{\t{int}}$ in \eqref{eq:H_int}, but without excited
 motional states, and with renormalized coupling constants in place of
@@
 
-Second, our result in \eqref{eq:H_2} neglects the effect of
+Second, the renormalization condition in \eqref{eq:counter_term}
+implies that the counter terms $\tilde G_X$ are second order in the
+coupling constants $G_X$, i.e.~$\tilde G_X\sim G_X^2$.  Although the
+effective Hamiltonian expansions in \eqref{eq:H_int_1_2} and
+\eqref{eq:H_int_3} are organized in powers of the interaction
+Hamiltonian $H_{\t{int}}$, the couplings $G_X$ are the ``small''
+parameters in which we can formally organize a perturbation theory;
+more specifically, the formally small quantities organizing our
+perturbation theory are two-body ground-state interaction energies
+(proportional to the couplings $G_X$) divided by the spectral gap of
+the single-atom Hamiltonian $H_0$.  If $M$ atoms can only couple
+through terms represented by a $p$-vertex diagrams for
+$p\ge p_M^{\t{min}}$, then the leading order contribution to $M$-body
+interactions is order $p_M^{\t{min}}$ in the couplings $G_X$.  If the
+same $p_M^{\t{min}}$-vertex diagrams involve any counter-terms,
+however, then these diagrams are at least order $p_M^{\t{min}}+1$ in
+the couplings $G_X$.  Counter-terms therefore only appear at
+next-to-leading order (NLO) in the calculation of effective $M$-body
+interactions.
+
+Finally, our result in \eqref{eq:H_2} neglects the effect of
 momentum-dependent two-body scattering.  When the effective range of
@@
 induced by terms proportional to
-$\c_{\mu q}^\dag \c_{\nu r}^\dag \c_{\sigma t} \c_{\rho s}$.  The
+$\c_{\rho s}^\dag \c_{\sigma t}^\dag \c_{\nu r} \c_{\mu q}$.  The
 minus sign in \eqref{eq:G_expanded} accounts for fermionic statistics:
@@
   \equiv  K \sum_{n>0} \f{K_n^2}{E_n^2},
+  \label{eq:alph_3_3}
 \end{align}
@@
 \end{align}
+Note that although \eqref{eq:H_3_3_S} is a loop diagram, the
+associated factors $\alpha_3^{(3)}$ and $\alpha_5^{(3)}$ in
+$H_3^{(3)}$ are finite.  At large $n$, atoms become free particles for
+which $n$ essentially indexes discrete momentum states.  These atoms
+thus have an energy which asymptotically scales as
+$E_n\sim n^2\equiv n_\x^2+n_\y^2+n_\z^2$.  Furthermore, the
+oscillatory behavior of atomic wavefunctions with increasing motional
+state indices $\ell,m$ implies that the overlap integral $K_{\ell m}$
+(and similarly $K^\ell_m$) becomes sharply peaked at $\ell\approx m$
+as $\ell$ and $m$ get large.  The asymptotic behavior of
+$\alpha_3^{(3)}$ at large $\ell,m,n$ is therefore
+\begin{align}
+  \alpha_3^{(3)}
+  \sim \int \f{\d^3\ell~\d^3m~\d^3n}{\p{\ell^2+m^2}\p{\ell^2+n^2}}
+  ~\delta\p{\ell-m}\delta\p{\ell-n}
+  \sim \int \f{\d^3\ell}{\ell^4}
+  \sim \int_{\ell_{\t{min}}}^\infty \f{\d\ell}{\ell^2}
+  \sim \f1{\ell_{\t{min}}},
+  \label{eq:alph_3_3_integral}
+\end{align}
+where in the last integral we changed to spherical coordinates, and
+$\ell_{\t{min}}^2$ is the minimum value of $\ell^2$ for which
+\begin{enumerate*}
+\item the energy $E_\ell\sim\ell^2$, and
+\item the integral expression in \eqref{eq:alph_3_3_integral} is a
+  good approximation to the corresponding sum in \eqref{eq:alph_3_3}.
+\end{enumerate*}
+Note that the introduction of $\ell_{\t{min}}$ amounts to neglecting a
+finite number of terms in the sum over $\ell,m,n$ in
+\eqref{eq:alph_3_3}, whose contribution to the value of
+$\alpha_3^{(3)}$ is finite.  Convergence of $\alpha_5^{(3)}$ is
+similarly guaranteed by the fact that the overlap integral $K_n$ does
+not asymptotically grow with increasing $n$, such that
+$\alpha_5^{(3)}$ asymptotically behaves as
+\begin{align}
+  \alpha_5^{(3)} \sim \int \f{\d^3 n}{n^4}
+  \sim \int_{n_{\t{min}}}^\infty \f{\d n}{n^2}
+  \sim \f1{n_{\t{min}}},
+  \label{eq:alph_5_3_integral}
+\end{align}
+where again $n_{\t{min}}$ is defined similarly to $\ell_{\t{min}}$.
 
@@
     single-site occupation numbers $N$ which are achievable in current
-    ${}^{87}$Sr experiments.}
-  \label{fig:coefficients}
+    ${}^{87}$Sr experiments.  \label{fig:coefficients}}
 \end{figure}
@@
     as a visual guide.  Resonance peaks are identified by the
-    many-body orbital states which are excited at the peak.}
-  \label{fig:sweep}
+    many-body orbital states which are excited at the peak.
+    \label{fig:sweep}}
 \end{figure}
@@
     states in the low-energy effective theory (see Appendix
-    \ref{sec:error}).}
-  \label{fig:shifts}
+    \ref{sec:error}).  \label{fig:shifts}}
 \end{figure}
@@
     in figure \ref{fig:shifts}, retrieved from
-    ref.~\cite{goban2018emergence}.}
-  \label{fig:shifts_summary}
+    ref.~\cite{goban2018emergence}.  \label{fig:shifts_summary}}
 \end{figure}
@@
     $\P_{NX}$, defined in \eqref{eq:P_NX}, is a projector onto the
-    space of the $N$-body orbital states $\ket{\N X}$.}
-  \label{fig:time}
+    space of the $N$-body orbital states $\ket{\N X}$.
+    \label{fig:time}}
 \end{figure}

### 2019-02-08 multi_body_interactions/papers/theory/reviews_NJP/revisions.tex
finished first draft of response to referee 1 from NJP

@@
 \usepackage[inline]{enumitem}
-\setlist[enumerate,1]{label={(\roman*)}}
+\setlist[enumerate]{leftmargin=*} % nice margins for enumerate
+
+% for feynman diagrams
+\usepackage{tikz,tikz-feynman}
+\tikzset{
+  baseline = (current bounding box.center)
+}
+\tikzfeynmanset{
+  compat = 1.1.0,
+  every feynman = {/tikzfeynman/small}
+}
+\newcommand{\shrink}[1]{\scalebox{0.8}{#1}} % for smaller diagrams
 
 % color definitions (used in a figure)
-\usepackage{xcolor} \newcommand{\blue}[1]{\textcolor{blue}{#1}}
+\usepackage{xcolor}
+\newcommand{\blue}[1]{\textcolor{blue}{#1}}
 \newcommand{\red}[1]{\textcolor{red}{#1}}
@@
   $k_L,k_L^x,k_L^y,k_L^z$ (denoting the wavenumber of the lattice
-  laser and its projection onto three principal axes) by
+  laser and its projection onto three lattice axes) by
   $k_{\text{L}}, k_{\text{L}}^\x, k_{\text{L}}^\y, k_{\text{L}}^\z$;
-  all $\U_x,\U_y,\U_z$ (denoting lattice depths along principal axes)
+  all $\U_x,\U_y,\U_z$ (denoting lattice depths along lattice axes)
   respectively by $\U_\x,\U_\y,\U_\z$; and all
@@
 
+\item We fixed an inconsistency in our convention for the ordering of
+  operators associated with the coupling symbol
+  $G^{\mu q;\nu r}_{\rho s;\sigma t}$.  The text which previously read
+  (page 14, line 47):
+
+  \red{for more general
+    $\p{\mu,q}+\p{\nu,r}\leftrightarrow\p{\rho,s}+\p{\sigma,t}$
+    coupling induced by terms proportional to
+    $\c_{\mu q}^\dag \c_{\nu r}^\dag \c_{\sigma t} \c_{\rho s}$.}
+
+  now reads:
+
+  \green{for more general
+    $\p{\mu,q}+\p{\nu,r}\leftrightarrow\p{\rho,s}+\p{\sigma,t}$
+    coupling induced by terms proportional to
+    $\c_{\rho s}^\dag \c_{\sigma t}^\dag \c_{\nu r} \c_{\mu q}$.}
+
+  Both examples are correct due to the fact that
+  $G^{\mu q;\nu r}_{\rho s;\sigma t}$ is self-adjoint, but the latter
+  ordering convention is used throughout the rest of the manuscript.
+
+\item In Eq.~(21), we have fixed an error in the nuclear spin indices
+  ($\mu,\nu$) on the two rightmost operators of the middle term, such
+  that (page 14, line 53):
+
+  \red{\begin{align*}
+      G^{\mu q;\nu r}_{\nu s;\mu t}
+      \c_{\nu s}^\dag \c_{\mu t}^\dag \c_{\nu r} \c_{\mu q}
+      = -G^{qr}_{ts} \c_{\nu s}^\dag \c_{\mu t}^\dag \c_{\mu r} \c_{\nu q}
+      = G^{qr}_{ts} \c_{\mu t}^\dag \c_{\nu s}^\dag \c_{\nu r} \c_{\mu q}.
+      \tag{21}
+    \end{align*}}
+
+  now reads:
+
+  \green{\begin{align*}
+      G^{\mu q;\nu r}_{\nu s;\mu t}
+      \c_{\nu s}^\dag \c_{\mu t}^\dag \c_{\nu r} \c_{\mu q}
+      = -G^{qr}_{ts} \c_{\nu s}^\dag \c_{\mu t}^\dag \c_{\nu r} \c_{\mu q}
+      = G^{qr}_{ts} \c_{\mu t}^\dag \c_{\nu s}^\dag \c_{\nu r} \c_{\mu q}.
+      \tag{21}
+    \end{align*}}
+
 \item To make our first appendix more self-contained, we have defined
-  $\P_0$ therein, such that the text (page 30, line 38):
+  the projector $\P_0$ used therein, such that the text (page 30, line
+  38):
 

### 2019-02-11 multi_body_interactions/papers/theory/reviews_NJP/referee_1_response.tex
merged some changes with AMR's suggestions

@@
   hope that thew revisions outlined below have made our manuscript
-  more accessible, and that we have addressed shortcomings identified
-  by the referee to a satisfactory degree.
+  more accessible, and that we have addressed the shortcomings
+  identified by the referee to a satisfactory degree.
 
@@
 
-  We thank the referee for pointing out the potential to mislead
+  We thank the referee for pointing out the possibility to mislead
   readers on the generality of our theory.  The referee is correct
@@
     experimental regime of (i) a deep lattice, with (ii) at most one
-    atom occupying each nuclear spin state on any lattice site, where
-    the latter restriction is a consequence of initial ground-state
+    atom occupying each nuclear spin state on any lattice site.  The
+    latter restriction is a consequence of initial ground-state
     preparation.  We fully characterize...}
 
-  We then added explicit mention of ground-state preparation to the
+  We now explicitly mention the ground-state preparation in the
   introduction, replacing (page 4, line 22):
@@
     occupying each nuclear spin state on any lattice site, which is a
-    consequence initial ground-state preparation.}
+    consequence of the experimental protocol which starts with all
+    atoms in the ground state.}
 
@@
     that have internal degrees of freedom and multiple collisional
-    parameters. We perform a direct comparison between the many-body
-    interaction energies predicted by our low-energy effective theory
-    and the experimental measurements of the density-dependent orbital
-    excitation spectra performed in ref.~[34], similarly to the
-    measurements with bosons performed in ref.~[38].  To facilitate
-    this comparison of excitation spectra and characterize the
-    low-lying excitations in our effective theory, we consider a
-    restriction of the theory to states with at most one orbital
-    excitation per lattice site, and find that the SU($N$) symmetry of
-    atomic collisions allow effective multi-body interactions to take
-    a remarkably simple form. We note that past work has also detected
-    experimental signatures of multi-body interactions in the form of
-    quantum phase revivals [39].}
-
-  We also clarify the limitations of our theory in the overview
-  between Eqs.~(8) and (9), as discussed later in point
-  \ref{pt:confusion}, and have made our technical contributions more
-  explicit in the preview of our main results (in section II),
-  replacing (page 7, line 48):
+    parameters.  Some past work has detected experimental signatures
+    of multi-body interactions in the form of quantum phase revivals
+    [39].  We instead compare the many-body interaction energies
+    predicted by our low-energy effective theory to the experimental
+    measurements of the density-dependent orbital excitation spectra
+    performed in ref.~[34], similarly to the measurements with bosons
+    performed in ref.~[38].  To facilitate this comparison of
+    excitation spectra and to characterize the low-lying excitations
+    in our effective theory, we consider a restriction of our theory
+    to states with at most one orbital excitation per lattice site.
+    In this case, we find that the SU($N$) symmetry of atomic
+    collisions allow the effective multi-body interactions to take a
+    remarkably simple form.}
+
+  We also clarify the limitations of our theory in the text between
+  Eqs.~(8) and (9), as discussed later in point \ref{pt:confusion},
+  and have made our technical contributions more explicit in the
+  description of our main results (in section II), replacing (page 7,
+  line 48):
 
@@
     where $H_M^{(p)}$ is an $M$-body Hamiltonian of order $p$ in the
-    coupling constants $G_X$, $I$ is the total nuclear spin of each
-    atom (e.g.~$I=9/2$ for ${}^{87}$Sr), and the sum terminates at
+    coupling constants $G_X$, and $I$ is the total nuclear spin of
+    each atom (e.g.~$I=9/2$ for ${}^{87}$Sr).  The sum terminates at
     $2I+1$ because this is the largest number of atoms which may
     initially occupy a single lattice site.  We explicitly compute all
-    $M$-body Hamiltonians $H_M\equiv\sum_p H_M^{(p)}$ through order
-    $p=3$, yielding effective two-, three-, and four-body
+    of the $M$-body Hamiltonians $H_M\equiv\sum_p H_M^{(p)}$ through
+    order $p=3$, yielding effective two-, three-, and four-body
     interactions.  To (i) facilitate a comparison with the
@@
     restriction of the multi-body Hamiltonians $H_M$ to states with at
-    most one orbital excitation per lattice site.  We find that the
-    SU($N$) symmetry of atomic collisions allows us to express
-    multi-body Hamiltonians in the simple form}
+    most one orbital excitation per lattice site.  Under this
+    restriction, we find that the SU($N$) symmetry of atomic
+    collisions allows us to express all multi-body Hamiltonians in the
+    simple form}
 
@@
     of at most one atom occupying each nuclear spin state on any
-    lattice site, we have derived a low-energy effective theory for
-    such atoms.  Our theory exhibits emergent multi-body interactions
-    that inherit the SU($N$) symmetry of atoms' bare, two-body
+    lattice site, we have derived a low-energy effective theory of
+    these atoms.  Our theory exhibits emergent multi-body interactions
+    that inherit the SU($N$) symmetry of the bare two-body
     interactions.  Considering a restriction of our theory to the
     subspace of at most one orbital excitation per lattice site, we
-    found that the SU($N$) symmetry of $M$-body Hamiltonians allowed
-    us to express them in a simple form, and to fully characterize
-    their eigenstates and spectra.}
+    found that the SU($N$) symmetry of all $M$-body Hamiltonians
+    allowed us to express them in a simple form, and to fully
+    characterize their eigenstates and spectra.}
 
@@
   particular highlighting that (i) our theory is limited to the case
-  of at most one atom per nuclear spin state per lattice site, and
+  of atoms in distinct nuclear spin states on each lattice site, and
   (ii) the theory in section III applies to any number of orbital
@@
   \green{The work in this paper is closely tied to the experimental
-    work in ref.~[34]; we begin with a short summary of the relevant
-    experimental procedures therein.  The experiment begins by
-    preparing a degenerate gas of $10^4$-$10^5$ (fermionic)
+    work reported in ref.~[34]; we begin with a short summary of the
+    relevant experimental procedures therein.  The experiment begins
+    by preparing a degenerate gas of $10^4$-$10^5$ (fermionic)
     ${}^{87}$Sr atoms in a uniform mixture of their ten nuclear spin
@@
     relevant to the experiment.  The temperature of the atoms is also
-    low enough to neglect thermal occupation of states outside the
-    ground-state manifold.}
-
-  \green{Once loaded into an optical lattice lattice, atoms are
-    addressed by an external (``clock'') interrogation laser with an
-    ultranarrow (26 mHz) linewidth, detuned by $\Delta$ from the
-    single-atom ${}^1S_0-{}^3P_0$ transition frequency $\omega_0$.
-    After a fixed interrogation time, the experiment turns off the
-    interrogation laser, removes all ground-state (${}^1S_0$) atoms
-    from the lattice, and uses absorption imaging to count any
-    remaining excited-state (${}^3P_0$) atoms.  Non-interacting atoms
-    in singly-occupied lattice sites undergo Rabi oscillations when
-    $\Delta=0$ during the interrogation time.  The orbital excitation
-    spectrum of multiply-occupied lattice sites, however, is shifted
-    by inter-atomic interactions, which results in spectroscopic peaks
+    low enough to neglect thermal occupation of motional states
+    outside the ground-state manifold.}
+
+  \green{Once loaded into an optical lattice, atoms are addressed by
+    an external (``clock'') interrogation laser with an ultranarrow
+    (26 mHz) linewidth, detuned by $\Delta$ from the single-atom
+    ${}^1S_0-{}^3P_0$ transition frequency $\omega_0$.  After a fixed
+    interrogation time, the experiment turns off the interrogation
+    laser, removes all ground-state (${}^1S_0$) atoms from the
+    lattice, and uses absorption imaging to count the remaining
+    excited-state (${}^3P_0$) atoms.  Non-interacting atoms in
+    singly-occupied lattice sites feature the typical single particle
+    lineshape peaked at $\Delta=0$.  The lineshapes of
+    multiply-occupied lattice sites, meanwhile, are shifted by
+    inter-atomic interactions, which results in spectroscopic peaks
     (i.e.~local maxima in excited-state atom counts) away from
@@
     that this spectroscopic protocol addresses only singly-excited
-    orbital states of any lattice site.  Doubly-excited states are
+    orbital states of lattice sites.  Doubly-excited states are
     manifestly off-resonant for single-photon absorption processes as
     $\abs{\Delta}\ll\omega_0$, while multi-photon absorption processes
-    are always off-resonant due to (i) the non-linearity ($\sim$kHz)
-    of orbital excitation energies for a collection of interacting
-    atoms, and (ii) the ultranarrow linewidth ($\sim$mHz) of the
-    interrogation laser.}
-
-  \green{Consider now a collection of fermionic atoms with total
-    nuclear spin $I$ and two atomic orbital states loaded into a
-    translationally invariant 3-D optical lattice.  While an external
-    trapping potential will generally break discrete translational
-    symmetry of the lattice, any background inhomogeneity can be made
-    negligible by spectroscopically addressing a sufficiently small
-    region of the lattice [34].  Throughout this paper, we work
-    strictly in the deep-lattice regime with negligible tunneling
-    between lattice sites, and assume that both atomic orbital states
-    experience identical lattice potentials.}
-
-  To reflect this change to section II, we have modified the
+    are always off-resonant due to (i) the interaction-induced
+    non-linearity ($\sim$kHz) of the orbital excitation energies, and
+    (ii) the ultranarrow linewidth ($\sim$mHz) of the interrogation
+    laser.}
+
+  \green{While an external trapping potential will generally break
+    discrete translational symmetry of the lattice, any background
+    inhomogeneity can be made negligible by spectroscopically
+    addressing a sufficiently small region of the lattice [34].
+    Throughout this paper, we work strictly in the deep-lattice regime
+    with negligible tunneling between lattice sites.  We also neglect
+    any lattice inhomogeneities and assume that both atomic orbital
+    states (i.e.~${}^1S_0$ and ${}^3P_0$) experience identical lattice
+    potentials.}
+
+  To reflect this change made to section II, we have modified the
   description of section II on page 4, line 52, previously:
@@
   \green{Current experiments with ${}^{87}$Sr can prepare up to five
-    atoms atoms in the same (ground) orbital state on a single lattice
-    site, and coherently address states with a single orbital
-    excitation per lattice site [34].  At ultracold temperatures well
-    below the single-atom motional excitation energies
-    $\Delta_n\equiv E_n-E_0$ for $n>0$, atoms only occupy their
-    motional ground state in the lattice.  For this reason, it is
-    common to map the description of these atoms onto a single-band
-    Hubbard model that captures all dynamics within the subspace of
-    motional ground states, i.e.~with single-atom wavefunctions
-    $\phi_{i,0}$.  When enough atoms occupy the same lattice site,
-    however, their interaction energy can approach the scale of the
-    motional excitation energies $\Delta_n$.  In this case, normally
-    far-off-resonant terms in the interaction Hamiltonian of (7) that
-    create atoms in excited motional states,
+    atoms in the same (ground) orbital state on a single lattice site,
+    and coherently address states with a single orbital excitation per
+    lattice site [34].  At ultracold temperatures well below the
+    single-atom motional excitation energies $\Delta_n\equiv E_n-E_0$
+    for $n>0$, atoms only occupy their motional ground state in the
+    lattice.  For this reason, it is common to map the description of
+    these atoms onto a single-band Hubbard model that captures all
+    dynamics within the subspace of motional ground states of
+    non-interacting atoms, i.e.~with single-atom wavefunctions
+    $\phi_{i,0}$.  Interactions, however, modify atoms' motional
+    ground-state wavefuctions.  The true motional ground state of a
+    collection of interacting atoms is then an admixture of the
+    non-interacting motional eigenstates, and a naive Hubbard model
+    that assumes atomic wavefunctions $\phi_{i,0}$ will fail to
+    reproduce the interacting atoms' orbital excitation spectrum.
+    Formally, corrections to the spectrum of interacting atoms can be
+    accounted for by a perturbative treatment of far-off-resonant
+    terms in the interaction Hamiltonian of (7) that create atoms in
+    excited motional states,
     e.g.~$\sim\c_{n\mu s}^\dag \c_{0,\nu t}^\dag \c_{0,\nu r}
-    \c_{0,\mu q}$ with $n>0$, become relevant.  The true motional
-    ground state of a collection of interacting atoms is then an
-    admixture of the non-interacting motional eigenstates, and a naive
-    Hubbard model that assumes atomic wavefunctions $\phi_{i,0}$ will
-    fail to reproduce the interacting atoms' orbital excitation
-    spectrum.}
-
-  \green{In order to understand interacting atoms' orbital excitation
-    spectrum in the context of a zero-temperature Hubbard-type model
-    that eliminates non-dynamical (i.e.~ground-state) motional degrees
-    of freedom, we develop a low-energy effective theory of
-    interacting AEAs in a deep lattice.  To simplify our theory, we
-    assume that any $N$ atoms on a single lattice site occupy $N$
-    nuclear spin states.  Ground-state preparation of the atoms in an
-    experiment implies that multiple occupation of a single nuclear
+    \c_{0,\mu q}$ with $n>0$.  These corrections can be understood
+    through interaction-induced {\it virtual} occupation of higher
+    bands (i.e.~excited motional states), which becomes relevant as
+    more atoms occupy the same lattice site, such that their
+    interaction energy becomes non-negligible compared to the motional
+    excitation energies $\Delta_n$.}
+
+  \green{In order to recast interaction-induced modifications to
+    orbital excitation spectra as corrections to the simple Hubbard
+    model (i.e.~computed using the non-interacting ground-state
+    wavefunctions $\phi_{i,0}$), we develop a low-energy effective
+    theory of interacting AEAs in a deep lattice.  To simplify our
+    theory, we assume that any $N$ atoms on a single lattice site
+    occupy distinct nuclear spin states.  This assumption applies for
+    any experimental protocol in which all atoms are initially
+    prepared in their orbital and motional ground states (as e.g.~in
+    ref.~[34]).  In this case, multiple occupation of a single nuclear
     spin state on any given lattice site is initially forbidden by
-    fermionic statistics, and subsequently cannot occur in the absence
-    of inter-site effects or hyperfine coupling between nuclear spin
-    states.}
+    fermionic statistics.  Subsequent violation of this condition
+    cannot occur in the absence of inter-site effects or hyperfine
+    coupling between nuclear spin states, as is the case in the
+    experiment in ref.~[34].}
 
@@
     where $H_2^{(\mu_1,\mu_2)}$ is a two-body Hamiltonian addressing
-    atoms with nuclear spin $\mu_1,\mu_2$;
+    atoms with nuclear spin $\mu_1,\mu_2$; and
     $\n_{\mu s}=\c_{\mu s}^\dag\c_{\mu s}$ is a number operator for
-    atoms with nuclear state $\mu$ and orbital state $s$; and the sum
-    is performed over all choices of nuclear spins $\mu_j$ with
+    atoms with nuclear state $\mu$ and orbital state $s$.  The sum in
+    (10) is performed over all choices of nuclear spins $\mu_j$ with
     $j=1,2,\cdots,M$ for which all $\mu_j$ are distinct, or
@@
   Note that the new equation number is 44 due to the addition of two
-  new equations in response to point \ref{pt:divergence} below.
+  new equations in response to point \ref{pt:divergence} below.  We
+  have made the same change to equation (F1), which is included in
+  Appendix F for quick reference.
 
@@
 
-  We thank the referee for their attention to detail, and for pointing
-  out the potential to mislead readers with respect to the nature of
-  $s$-wave interactions by our expression of Eq.~(2).  We have
-  therefore added the a paragraph to section II, following the
+  We thank the referee for their attention to this issue, and for
+  pointing out the potential to mislead readers with respect to the
+  nature of $s$-wave interactions by our expression of Eq.~(2).  We
+  have therefore added the a paragraph to section II, following the
   re-statement of Eq.~(2) in Eq.~(5).  Specifically, the following
@@
     and separately accounting for momentum-dependent scattering is
-    essentially the same as that used for similar calculations in
-    refs.~[35--37].  While this approach does not provide insight into
-    the microscopic nature of atomic interactions, it is suitable for
-    the phenomenological description of these interactions, and
+    essentially the same as the approach used for similar calculations
+    in refs.~[35--37].  While this approach does not provide insight
+    into the microscopic nature of atomic interactions, it is suitable
+    for the phenomenological description of these interactions, and
     particular for our eventual development of a low-energy effective
@@
 
-  \green{On physical grounds, the net effective two-body interaction
-    must clearly be finite, but individual sums over the excited
-    states in loop diagrams of (15) may generally diverge [35].  These
+  \green{On physical grounds, the net two-body interaction must
+    clearly be finite, but individual sums over the excited states in
+    loop diagrams of (15) may generally diverge [35].  These
     divergences ultimately appear due to our use of effective field
@@
     spectrum at zero temperature, in principle we need to identify
-    these motional ground states.  We can then ignore all excited
+    this motional ground state.  We can then ignore all excited
     motional states, which will be neither thermally occupied nor
@@
     interactions are sufficiently weak compared to the spectral gap
-    $\Delta$ of the non-interacting Hamiltonian $H_0$ between
-    $\H_{\t{ground}}^{\t{single}}$ and its orthogonal complement
-    $\H_{\t{full}}\setminus\H_{\t{ground}}^{\t{single}}$, one can
-    identify a particular unitary operator $U$ (acting on the full
-    Hilbert space $\H_{\t{full}}$) which rotates
+    $\Delta$ between $\H_{\t{ground}}^{\t{single}}$ and its orthogonal
+    complement $\H_{\t{full}}\setminus\H_{\t{ground}}^{\t{single}}$,
+    one can identify a particular unitary operator $U$ (acting on the
+    full Hilbert space $\H_{\t{full}}$) which rotates
     $\H_{\t{ground}}^{\t{multi}}$ into $\H_{\t{ground}}^{\t{single}}$
@@
     $\H_{\t{ground}}^{\t{multi}}$.  The use of an effective
-    Hamiltonian $H_{\t{eff}}$ thus side-steps the need to identify
+    Hamiltonian $H_{\t{eff}}$ thus overcomes the need to identify
     $\H_{\t{ground}}^{\t{multi}}$ in order to compute the orbital
@@
     motional state indices $\ell,m$ implies that the overlap integral
-    $K_{\ell m}$ (and similarly $K^\ell_m$) becomes sharply peaked at
-    $\ell\approx m$ as $\ell$ and $m$ get large.  The asymptotic
-    behavior of $\alpha_3^{(3)}$ at large $\ell,m,n$ is therefore
+    $K_{\ell m}$ becomes sharply peaked at $\ell\approx m$ as $\ell$
+    and $m$ get large.  The asymptotic behavior of $\alpha_3^{(3)}$ at
+    large $\ell,m,n$ is therefore
     \begin{align*}
@@
   We thank the referee for pointing out that our manuscript would
-  benefit from a careful explanation of our renormalization scheme.
-  Renormalization is a class of techniques, the details of which may
-  look different in the context of different field theories; worse
-  still, any chosen means to renormalize a field theory is generally
-  not unique.
+  benefit from a more careful explanation of our renormalization
+  scheme.  Renormalization is a technique whose details may look
+  different in the context of different field theories; worse still,
+  any chosen means to renormalize a field theory is generally not
+  unique.
 
@@
     $\tilde G_X \equiv G_X^{\t{bare}} - G_X$.  For convenience, we can
-    choose the values of $G_X$ to be those of the physically observed
-    effective coupling constants on the right-hand side of (15).
-    Representing the new coupling constants $G_X$ by regular vertices
-    and the counter-terms $\tilde G_X$ by a crossed dot
-    (i.e.~$\otimes$), this choice leads to the {\it renormalization
-      condition}
+    choose the values of $G_X$ to be those of the net effective
+    coupling constants on the right-hand side of (15).  Representing
+    the new coupling constants $G_X$ by regular vertices and the
+    counter-terms $\tilde G_X$ by a crossed dot (i.e.~$\otimes$), this
+    choice leads to the {\it renormalization condition}
     \begin{align*}
@@
     This renormalization condition has the benefit of allowing us to
-    express effective two-body interactions simply in terms of
-    physically observed effective two-body coupling constants, rather
-    than in terms of long sums at all order of the bare coupling
-    constants.}
+    express effective two-body interactions simply in terms of net
+    effective two-body coupling constants, rather than in terms of
+    long sums at all order of the bare coupling constants.}
+
+  We have further clarified the implications of our renormalization
+  condition in the discussion following Eq.~(17), replacing the
+  following text (page 13, line 17):
+
+  \red{Instead, we must first compute the effective coupling constants
+    $G_X^{\t{lattice}}\p{\U}$ in a lattice, which now depend on the
+    lattice depth $\U$, and in turn use the effective coupling
+    constants to compute interaction energies.}
+
+  by:
+
+  \green{The renormalization condition in (16) explicitly fixes $G_X$
+    to the net effective coupling constants in any given setting.
+    Instead of using free-space coupling constants to compute
+    interaction energies in a lattice, we must therefore first compute
+    the effective coupling constants $G_X^{\t{lattice}}\p{\U}$, which
+    now depend on the lattice depth $\U$, and in turn use these
+    effective coupling constants to compute interaction energies.}
 

### 2019-02-11 multi_body_interactions/papers/theory/paper.tex
merged some changes with AMR's suggestions

@@
 
-\usepackage{setspace} % to modify spacing in table of contents
-
 % leave a note in the text, visible in the compiled document
@@
   \item at most one atom occupying each nuclear spin state on any
-    lattice site,
+    lattice site.
   \end{enumerate*}
-  where the latter restriction is a consequence of initial
-  ground-state preparation.  We fully characterize the low-lying
-  excitations in our effective theory, and compare predictions of
-  many-body interaction energies with direct measurements of many-body
-  excitation spectra in an optical lattice clock.  Our work makes the
-  first step in enabling a controlled, bottom-up experimental
-  investigation of multi-body SU($N$) physics.
+  The latter restriction is a consequence of initial ground-state
+  preparation.  We fully characterize the low-lying excitations in our
+  effective theory, and compare predictions of many-body interaction
+  energies with direct measurements of many-body excitation spectra in
+  an optical lattice clock.  Our work makes the first step in enabling
+  a controlled, bottom-up experimental investigation of multi-body
+  SU($N$) physics.
 \end{abstract}
@@
 
-\begin{spacing}{1.3}
 \tableofcontents
-\end{spacing}
 
@@
 to the experimental regime of at most one atom occupying each nuclear
-spin state on any lattice site, which is a consequence initial
-ground-state preparation.
+spin state on any lattice site, which is a consequence of the
+experimental protocol which starts with all atoms in the ground state.
 
@@
 for the first time with fermions that have internal degrees of freedom
-and multiple collisional parameters.  We perform a direct comparison
-between the many-body interaction energies predicted by our low-energy
-effective theory and the experimental measurements of the
+and multiple collisional parameters.  Some past work has detected
+experimental signatures of multi-body interactions in the form of
+quantum phase revivals \cite{will2010timeresolved}.  We instead
+compare the many-body interaction energies predicted by our low-energy
+effective theory to the experimental measurements of the
 density-dependent orbital excitation spectra performed in
@@
 bosons performed in ref.~\cite{mark2011precision}.  To facilitate this
-comparison of excitation spectra and characterize the low-lying
-excitations in our effective theory, we consider a restriction of the
-theory to states with at most one orbital excitation per lattice site,
-and find that the SU($N$) symmetry of atomic collisions allow
-effective multi-body interactions to take a remarkably simple form.
-We note that past work has also detected experimental signatures of
-multi-body interactions in the form of quantum phase revivals
-\cite{will2010timeresolved}.
+comparison of excitation spectra and to characterize the low-lying
+excitations in our effective theory, we consider a restriction of our
+theory to states with at most one orbital excitation per lattice site.
+In this case, we find that the SU($N$) symmetry of atomic collisions
+allow the effective multi-body interactions to take a remarkably
+simple form.
 
@@
 
-The work in this paper is closely tied to the experimental work in
-ref.~\cite{goban2018emergence}; we begin with a short summary of the
-relevant experimental procedures therein.  The experiment begins by
-preparing a degenerate gas of $10^4$-$10^5$ (fermionic) ${}^{87}$Sr
-atoms in a uniform mixture of their ten nuclear spin states and at
-$\sim0.1$ of their fermi temperature ($\sim10$ nanokelvin)
-\cite{campbell2017fermidegenerate, marti2018imaging}.  This gas is
-loaded into a primitive cubic optical lattice at the ``magic
-wavelength'' for which both ground (${}^1S_0$) and first-excited
-(${}^3P_0$) electronic orbital states of the atoms experience the same
-lattice potential \cite{ye2008quantum}.  Lattice depths along the
-principal axes of the lattice are roughly equal in magnitude, with a
-geometric mean that can be varied from 30 to 80 $E_{\t{R}}$, where
-$E_{\t{R}}\approx3.5\times2\pi~\t{kHz}$ is the lattice photon recoil
-energy of the atoms (with the reduced Planck constant $\hbar=1$
-throughout this paper).  These lattice depths are sufficiently large
-as to neglect tunneling on the time scales relevant to the experiment.
-The temperature of the atoms is also low enough to neglect thermal
-occupation of states outside the ground-state manifold.
-
-Once loaded into an optical lattice lattice, atoms are addressed by an
+The work in this paper is closely tied to the experimental work
+reported in ref.~\cite{goban2018emergence}; we begin with a short
+summary of the relevant experimental procedures therein.  The
+experiment begins by preparing a degenerate gas of $10^4$-$10^5$
+(fermionic) ${}^{87}$Sr atoms in a uniform mixture of their ten
+nuclear spin states and at $\sim0.1$ of their fermi temperature
+($\sim10$ nanokelvin) \cite{campbell2017fermidegenerate,
+  marti2018imaging}.  This gas is loaded into a primitive cubic
+optical lattice at the ``magic wavelength'' for which both ground
+(${}^1S_0$) and first-excited (${}^3P_0$) electronic orbital states of
+the atoms experience the same lattice potential \cite{ye2008quantum}.
+Lattice depths along the principal axes of the lattice are roughly
+equal in magnitude, with a geometric mean that can be varied from 30
+to 80 $E_{\t{R}}$, where $E_{\t{R}}\approx3.5\times2\pi~\t{kHz}$ is
+the lattice photon recoil energy of the atoms (with the reduced Planck
+constant $\hbar=1$ throughout this paper).  These lattice depths are
+sufficiently large as to neglect tunneling on the time scales relevant
+to the experiment.  The temperature of the atoms is also low enough to
+neglect thermal occupation of motional states outside the ground-state
+manifold.
+
+Once loaded into an optical lattice, atoms are addressed by an
 external (``clock'') interrogation laser with an ultranarrow (26 mHz)
@@
 ground-state (${}^1S_0$) atoms from the lattice, and uses absorption
-imaging to count any remaining excited-state (${}^3P_0$) atoms.
-Non-interacting atoms in singly-occupied lattice sites undergo Rabi
-oscillations when $\Delta=0$ during the interrogation time.  The
-orbital excitation spectrum of multiply-occupied lattice sites,
-however, is shifted by inter-atomic interactions, which results in
-spectroscopic peaks (i.e.~local maxima in excited-state atom counts)
-away from $\Delta=0$.  A sweep across different detunings $\Delta$ (on
-the scale of inter-atomic interaction energies) thus constitutes a
-measurement of the many-body orbital excitation spectrum.  We note
-that this spectroscopic protocol addresses only singly-excited orbital
-states of any lattice site.  Doubly-excited states are manifestly
-off-resonant for single-photon absorption processes as
-$\abs{\Delta}\ll\omega_0$, while multi-photon absorption processes are
-always off-resonant due to
+imaging to count the remaining excited-state (${}^3P_0$) atoms.
+Non-interacting atoms in singly-occupied lattice sites feature the
+typical single particle lineshape peaked at $\Delta=0$.  The
+lineshapes of multiply-occupied lattice sites, meanwhile, are shifted
+by inter-atomic interactions, which results in spectroscopic peaks
+(i.e.~local maxima in excited-state atom counts) away from $\Delta=0$.
+A sweep across different detunings $\Delta$ (on the scale of
+inter-atomic interaction energies) thus constitutes a measurement of
+the many-body orbital excitation spectrum.  We note that this
+spectroscopic protocol addresses only singly-excited orbital states of
+lattice sites.  Doubly-excited states are manifestly off-resonant for
+single-photon absorption processes as $\abs{\Delta}\ll\omega_0$, while
+multi-photon absorption processes are always off-resonant due to
 \begin{enumerate*}
-\item the non-linearity ($\sim$kHz) of orbital excitation energies for
-  a collection of interacting atoms, and
+\item the interaction-induced non-linearity ($\sim$kHz) of the orbital
+  excitation energies, and
 \item the ultranarrow linewidth ($\sim$mHz) of the interrogation
@@
 
-Consider now a collection of fermionic atoms with total nuclear spin
-$I$ and two atomic orbital states loaded into a translationally
-invariant 3-D optical lattice.  While an external trapping potential
-will generally break discrete translational symmetry of the lattice,
-any background inhomogeneity can be made negligible by
-spectroscopically addressing a sufficiently small region of the
-lattice \cite{goban2018emergence}.  Throughout this paper, we work
-strictly in the deep-lattice regime with negligible tunneling between
-lattice sites, and assume that both atomic orbital states experience
-identical lattice potentials.  The single-particle Hamiltonian of the
-atoms can then be written in the form
+While an external trapping potential will generally break discrete
+translational symmetry of the lattice, any background inhomogeneity
+can be made negligible by spectroscopically addressing a sufficiently
+small region of the lattice \cite{goban2018emergence}.  Throughout
+this paper, we work strictly in the deep-lattice regime with
+negligible tunneling between lattice sites.  We also neglect any
+lattice inhomogeneities and assume that both atomic orbital states
+(i.e.~${}^1S_0$ and ${}^3P_0$) experience identical lattice
+potentials.  The single-particle Hamiltonian of the atoms can then be
+written in the form
 \begin{align}
@@
 separately accounting for momentum-dependent scattering is essentially
-the same as that used for similar calculations in
+the same as the approach used for similar calculations in
 refs.~\cite{johnson2012effective, yin2014universal,
@@
 
-Current experiments with ${}^{87}$Sr can prepare up to five atoms
-atoms in the same (ground) orbital state on a single lattice site, and
+Current experiments with ${}^{87}$Sr can prepare up to five atoms in
+the same (ground) orbital state on a single lattice site, and
 coherently address states with a single orbital excitation per lattice
@@
 these atoms onto a single-band Hubbard model that captures all
-dynamics within the subspace of motional ground states, i.e.~with
-single-atom wavefunctions $\phi_{i,0}$.  When enough atoms occupy the
-same lattice site, however, their interaction energy can approach the
-scale of the motional excitation energies $\Delta_n$.  In this case,
-normally far-off-resonant terms in the interaction Hamiltonian of
-\eqref{eq:H_int} that create atoms in excited motional states,
-e.g.~$\sim\c_{n\mu s}^\dag \c_{0,\nu t}^\dag \c_{0,\nu r} \c_{0,\mu
-  q}$ with $n>0$, become relevant.  The true motional ground state of
-a collection of interacting atoms is then an admixture of the
+dynamics within the subspace of motional ground states of
+non-interacting atoms, i.e.~with single-atom wavefunctions
+$\phi_{i,0}$.  Interactions, however, modify atoms' motional
+ground-state wavefuctions.  The true motional ground state of a
+collection of interacting atoms is then an admixture of the
 non-interacting motional eigenstates, and a naive Hubbard model that
 assumes atomic wavefunctions $\phi_{i,0}$ will fail to reproduce the
-interacting atoms' orbital excitation spectrum.
-
-In order to understand interacting atoms' orbital excitation spectrum
-in the context of a zero-temperature Hubbard-type model that
-eliminates non-dynamical (i.e.~ground-state) motional degrees of
-freedom, we develop a low-energy effective theory of interacting AEAs
-in a deep lattice.  To simplify our theory, we assume that any $N$
-atoms on a single lattice site occupy $N$ nuclear spin states.
-Ground-state preparation of the atoms in an experiment implies that
-multiple occupation of a single nuclear spin state on any given
-lattice site is initially forbidden by fermionic statistics, and
-subsequently cannot occur in the absence of inter-site effects or
-hyperfine coupling between nuclear spin states.
+interacting atoms' orbital excitation spectrum.  Formally, corrections
+to the spectrum of interacting atoms can be accounted for by a
+perturbative treatment of far-off-resonant terms in the interaction
+Hamiltonian of \eqref{eq:H_int} that create atoms in excited motional
+states,
+e.g.~$\sim\c_{n\mu s}^\dag \c_{0,\nu t}^\dag \c_{0,\nu r} \c_{0,\mu
+  q}$ with $n>0$.  These corrections can be understood through
+interaction-induced {\it virtual} occupation of higher bands
+(i.e.~excited motional states), which becomes relevant as more atoms
+occupy the same lattice site, such that their interaction energy
+becomes non-negligible compared to the motional excitation energies
+$\Delta_n$.
+
+In order to recast interaction-induced modifications to orbital
+excitation spectra as corrections to the simple Hubbard model
+(i.e.~computed using the non-interacting ground-state wavefunctions
+$\phi_{i,0}$), we develop a low-energy effective theory of interacting
+AEAs in a deep lattice.  To simplify our theory, we assume that any
+$N$ atoms on a single lattice site occupy distinct nuclear spin
+states.  This assumption applies for any experimental protocol in
+which all atoms are initially prepared in their orbital and motional
+ground states (as e.g.~in ref.~\cite{goban2018emergence}).  In this
+case, multiple occupation of a single nuclear spin state on any given
+lattice site is initially forbidden by fermionic statistics.
+Subsequent violation of this condition cannot occur in the absence of
+inter-site effects or hyperfine coupling between nuclear spin states,
+as is the case of the experiment in ref.~\cite{goban2018emergence}.
 
@@
 where $H_M^{(p)}$ is an $M$-body Hamiltonian of order $p$ in the
-coupling constants $G_X$, $I$ is the total nuclear spin of each atom
-(e.g.~$I=9/2$ for ${}^{87}$Sr), and the sum terminates at $2I+1$
+coupling constants $G_X$, and $I$ is the total nuclear spin of each
+atom (e.g.~$I=9/2$ for ${}^{87}$Sr).  The sum terminates at $2I+1$
 because this is the largest number of atoms which may initially occupy
@@
   many-body orbital excitation spectra performed in
-  ref.~\cite{goban2018emergence} and
+  ref.~\cite{goban2018emergence} andr
 \item characterize the low-lying excitations in our effective theory,
@@
 this restriction, we find that the SU($N$) symmetry of atomic
-collisions allows us to express multi-body Hamiltonians in the simple
-form
+collisions allows us to express all multi-body Hamiltonians in the
+simple form
 \begin{align}
@@
 where $H_2^{(\mu_1,\mu_2)}$ is a two-body Hamiltonian addressing atoms
-with nuclear spin $\mu_1,\mu_2$;
+with nuclear spin $\mu_1,\mu_2$; and
 $\n_{\mu s}=\c_{\mu s}^\dag\c_{\mu s}$ is a number operator for atoms
-with nuclear state $\mu$ and orbital state $s$; and the sum is
-performed over all choices of nuclear spins $\mu_j$ with
-$j=1,2,\cdots,M$ for which all $\mu_j$ are distinct, or equivalently
-all choices of $\mu_j$ for which the set $\set{\mu_j}$ contains $M$
-elements, for a total of ${2I+1\choose M}\times\p{M!}$ nuclear spin
-combinations.  The key feature of the $M$-body interactions in
-\eqref{eq:H_M_preview} is that they ultimately take the same form as
-two-body interactions, but with the addition of $M-2$ spectator atoms.
-This form is a direct consequence of the SU($N$) symmetry of
-underlying two-body interactions.
+with nuclear state $\mu$ and orbital state $s$.  The sum in
+\eqref{eq:H_M_preview} is performed over all choices of nuclear spins
+$\mu_j$ with $j=1,2,\cdots,M$ for which all $\mu_j$ are distinct, or
+equivalently all choices of $\mu_j$ for which the set $\set{\mu_j}$
+contains $M$ elements, for a total of ${2I+1\choose M}\times\p{M!}$
+nuclear spin combinations.  The key feature of the $M$-body
+interactions in \eqref{eq:H_M_preview} is that they ultimately take
+the same form as two-body interactions, but with the addition of $M-2$
+spectator atoms.  This form is a direct consequence of the SU($N$)
+symmetry of underlying two-body interactions.
 
@@
 an orbital excitation spectrum at zero temperature, in principle we
-need to identify these motional ground states.  We can then ignore all
+need to identify this motional ground state.  We can then ignore all
 excited motional states, which will be neither thermally occupied nor
@@
 interactions are sufficiently weak compared to the spectral gap
-$\Delta$ of the non-interacting Hamiltonian $H_0$ between
-$\H_{\t{ground}}^{\t{single}}$ and its orthogonal complement
-$\H_{\t{full}}\setminus\H_{\t{ground}}^{\t{single}}$, one can identify
-a particular unitary operator $U$ (acting on the full Hilbert space
-$\H_{\t{full}}$) which rotates $\H_{\t{ground}}^{\t{multi}}$ into
-$\H_{\t{ground}}^{\t{single}}$ \cite{bravyi2011schrieffer}.  This
-unitary $U$ can be used to construct an {\it effective Hamiltonian}
-$H_{\t{eff}} = U H U^\dag$ with two key properties:
+$\Delta$ between $\H_{\t{ground}}^{\t{single}}$ and its orthogonal
+complement $\H_{\t{full}}\setminus\H_{\t{ground}}^{\t{single}}$, one
+can identify a particular unitary operator $U$ (acting on the full
+Hilbert space $\H_{\t{full}}$) which rotates
+$\H_{\t{ground}}^{\t{multi}}$ into $\H_{\t{ground}}^{\t{single}}$
+\cite{bravyi2011schrieffer}.  This unitary $U$ can be used to
+construct an {\it effective Hamiltonian} $H_{\t{eff}} = U H U^\dag$
+with two key properties:
 \begin{enumerate*}
@@
 \end{enumerate*}
-The use of an effective Hamiltonian $H_{\t{eff}}$ thus side-steps the
+The use of an effective Hamiltonian $H_{\t{eff}}$ thus overcomes the
 need to identify $\H_{\t{ground}}^{\t{multi}}$ in order to compute the
@@
 where the blob on the right schematically represents the net effective
-two-body interaction.  On physical grounds, the net effective two-body
+two-body interaction.  On physical grounds, the net two-body
 interaction must clearly be finite, but individual sums over excited
@@
 $\tilde G_X \equiv G_X^{\t{bare}} - G_X$.  For convenience, we can
-choose the values of $G_X$ to be those of the physically observed
-effective coupling constants on the right-hand side of
-\eqref{eq:H_2_expansion}.  Representing the new coupling constants
-$G_X$ by regular vertices and the counter-terms $\tilde G_X$ by a
-crossed dot (i.e.~$\otimes$), this choice leads to the {\it
-  renormalization condition}
+choose the values of $G_X$ to be those of the net effective coupling
+constants on the right-hand side of \eqref{eq:H_2_expansion}.
+Representing the new coupling constants $G_X$ by regular vertices and
+the counter-terms $\tilde G_X$ by a crossed dot (i.e.~$\otimes$), this
+choice leads to the {\it renormalization condition}
 \begin{align}
@@
 This renormalization condition has the benefit of allowing us to
-express effective two-body interactions simply in terms of physically
-observed effective two-body coupling constants, rather than in terms
-of long sums at all order of the bare coupling constants.  By
-construction, the counter-terms we have introduced exactly cancel all
-terms beyond leading order in \eqref{eq:H_2_expansion}, which implies
-that the effective two-body interaction Hamiltonian is simply
+express effective two-body interactions simply in terms of net
+effective two-body coupling constants, rather than in terms of long
+sums at all order of the bare coupling constants.  By construction,
+the counter-terms we have introduced exactly cancel all terms beyond
+leading order in \eqref{eq:H_2_expansion}, which implies that the
+effective two-body interaction Hamiltonian is simply
 \begin{align}
@@
 are a few comments we must make concerning renormalization and the
-result in \eqref{eq:H_2}.  First, this effective two-body interaction
+result in \eqref{eq:H_2}.  First, the effective two-body interaction
 $H_2$ in \eqref{eq:H_2} takes the same form as the bare two-body
 interaction $H_{\t{int}}$ in \eqref{eq:H_int}, but without excited
-motional states, and with renormalized coupling constants in place of
-bare ones.  Our choice of renormalization scheme, while convenient for
-the analytical development of a low-energy effective theory, no longer
-allows us to use the coupling constants $G_X$ as defined by the
-free-space scattering lengths $a_X$ in \eqref{eq:couplings} to compute
-interaction energies.  Instead, we must first compute the effective
-coupling constants $G_X^{\t{lattice}}\p{\U}$ in a lattice, which now
-depend on the lattice depth $\U$, and in turn use the effective
-coupling constants to compute interaction energies.  We discuss the
-calculation of effective coupling constants in Appendix
-\ref{sec:renormalization}.
+motional states, and with renormalized coupling constants.  Our choice
+of renormalization scheme, while convenient for the analytical
+development of a low-energy effective theory, no longer allows us to
+use the coupling constants $G_X$ as defined by the free-space
+scattering lengths $a_X$ in \eqref{eq:couplings} to compute
+interaction energies.  The renormalization condition in
+\eqref{eq:counter_term} explicitly fixes $G_X$ to the net effective
+coupling constants in any given setting.  Instead of using free-space
+coupling constants to compute interaction energies in a lattice, we
+must therefore first compute the effective coupling constants
+$G_X^{\t{lattice}}\p{\U}$, which now depend on the lattice depth $\U$,
+and in turn use these effective coupling constants to compute
+interaction energies.  We discuss the calculation of effective
+coupling constants in Appendix \ref{sec:renormalization}.
 
@@
 state indices $\ell,m$ implies that the overlap integral $K_{\ell m}$
-(and similarly $K^\ell_m$) becomes sharply peaked at $\ell\approx m$
-as $\ell$ and $m$ get large.  The asymptotic behavior of
-$\alpha_3^{(3)}$ at large $\ell,m,n$ is therefore
+becomes sharply peaked at $\ell\approx m$ as $\ell$ and $m$ get large.
+The asymptotic behavior of $\alpha_3^{(3)}$ at large $\ell,m,n$ is
+therefore
 \begin{align}
@@
 regime of at most one atom occupying each nuclear spin state on any
-lattice site, we have derived a low-energy effective theory for such
+lattice site, we have derived a low-energy effective theory of these
 atoms.  Our theory exhibits emergent multi-body interactions that
-inherit the SU($N$) symmetry of atoms' bare, two-body interactions.
+inherit the SU($N$) symmetry of the bare two-body interactions.
 Considering a restriction of our theory to the subspace of at most one
 orbital excitation per lattice site, we found that the SU($N$)
-symmetry of $M$-body Hamiltonians allowed us to express them in a
+symmetry of all $M$-body Hamiltonians allowed us to express them in a
 simple form, and to fully characterize their eigenstates and spectra.
@@
   H_M = \sum_{\abs{\set{\mu_j}}=M}
-  \p{U_{M,\g} \prod_{\alpha=1}^M \n_{\mu_\alpha,\g}
-    + U_{M,+} \n_{\mu_1,\e} \prod_{\alpha=2}^M \n_{\mu_\alpha,\g}
+  \p{U_{M,\g} \n_{\mu_1,\g} \n_{\mu_2,\g}
+    + U_{M,+} \n_{\mu_1,\e} \n_{\mu_2,\g}
     + U_{M,-} \c_{\mu_1,\g}^\dag \c_{\mu_2,\e}^\dag
-    \c_{\mu_2,\g} \c_{\mu_1,\e} \prod_{\alpha=3}^M \n_{\mu_\alpha,\g}},
+    \c_{\mu_2,\g} \c_{\mu_1,\e}}
+  \prod_{\alpha=3}^M \n_{\mu_\alpha,\g},
 \end{align}

### 2019-02-11 multi_body_interactions/papers/theory/reviews_NJP/referee_1_response.tex
final changes to referee 1 response (NJP)

@@
 Excerpts from the referee reports are written in \blue{blue}, excerpts
-from our original manuscript are written in \red{red}, and new text in
-the revised version of our manuscript is written in \green{green}.
-All page, line, and reference numbers generally refer to those in our
-original manuscript, unless explicitly stated otherwise.
+from our original manuscript are written in \red{red}, and excerpts
+from the revised version of our manuscript is written in
+\green{green}.  All page, line, equation, and reference numbers
+generally refer to those in our original manuscript, unless explicitly
+stated otherwise.
 
@@
     performed in ref.~[34] and (ii) characterize the low-lying
-    excitations in our effective theory, we also consider a
-    restriction of the multi-body Hamiltonians $H_M$ to states with at
-    most one orbital excitation per lattice site.  Under this
-    restriction, we find that the SU($N$) symmetry of atomic
-    collisions allows us to express all multi-body Hamiltonians in the
-    simple form}
+    excitations in our effective theory, we additionally restrict the
+    multi-body Hamiltonians $H_M$ to states with at most one orbital
+    excitation per lattice site.  Under this restriction, we find that
+    the SU($N$) symmetry of atomic collisions allows us to express all
+    multi-body Hamiltonians in the simple form}
 
@@
     orbital states of lattice sites.  Doubly-excited states are
-    manifestly off-resonant for single-photon absorption processes as
-    $\abs{\Delta}\ll\omega_0$, while multi-photon absorption processes
-    are always off-resonant due to (i) the interaction-induced
-    non-linearity ($\sim$kHz) of the orbital excitation energies, and
-    (ii) the ultranarrow linewidth ($\sim$mHz) of the interrogation
-    laser.}
+    off-resonant due to (i) the interaction-induced non-linearity
+    ($\sim$kHz) of orbital excitation energies, and (ii) the
+    ultranarrow linewidth ($\sim$mHz) of the interrogation laser.}
 

### 2019-02-11 multi_body_interactions/papers/theory/paper.tex
final changes to referee 1 response (NJP)

@@
 spectroscopic protocol addresses only singly-excited orbital states of
-lattice sites.  Doubly-excited states are manifestly off-resonant for
-single-photon absorption processes as $\abs{\Delta}\ll\omega_0$, while
-multi-photon absorption processes are always off-resonant due to
+lattice sites.  Doubly-excited states are off resonant due to
 \begin{enumerate*}
@@
   many-body orbital excitation spectra performed in
-  ref.~\cite{goban2018emergence} andr
+  ref.~\cite{goban2018emergence} and
 \item characterize the low-lying excitations in our effective theory,
 \end{enumerate*}
-we also consider a restriction of the multi-body Hamiltonians $H_M$ to
-states with at most one orbital excitation per lattice site.  Under
-this restriction, we find that the SU($N$) symmetry of atomic
-collisions allows us to express all multi-body Hamiltonians in the
-simple form
+we additionally restrict the multi-body Hamiltonians $H_M$ to states
+with at most one orbital excitation per lattice site.  Under this
+restriction, we find that the SU($N$) symmetry of atomic collisions
+allows us to express all multi-body Hamiltonians in the simple form
 \begin{align}
@@
 order term, the expansion involves {\it effective corrections} to the
-action of $H_{\t{int}}$ on the single-particle motional ground states
+action of $H_{\t{int}}$ on the non-interacting motional ground states
 (i.e.~on $\H_{\t{ground}}^{\t{single}}$) in the form of higher-order
@@
 \cite{johnson2012effective} we define $\alpha_2^{(1)} \equiv K$ as the
-overlap integral between two atoms occupying single-particle motional
+overlap integral between two atoms occupying non-interacting motional
 ground states.
@@
 $\ket{\N X}$.  Interrogating the atoms for a time $t$ by a laser
-resonant with the excitation energy $\Delta_{N,\pm}$ then gives us the
+resonant with the excitation energy $\Delta_{N\pm}$ then gives us the
 state
@@
   \rho_N^{(\pm)}\p{t} = \f1{\Tr\P_{N,0}}
-  \sum_{\abs{\N}=N} \exp\p{-it\Omega_{\N,\pm}S_{\N,\pm}^\x}
-  \op{\N,0} \exp\p{it\Omega_{\N,\pm}S_{\N,\pm}^\x},
+  \sum_{\abs{\N}=N} \exp\p{-it\Omega_{\N\pm}S_{\N\pm}^\x}
+  \op{\N,0} \exp\p{it\Omega_{\N\pm}S_{\N\pm}^\x},
 \end{align}
 where the Rabi frequencies $\Omega_{\N,+},\Omega_{\N,-}$ and
-pseudo-spin operators $S_{\N,\pm}^\x$ are respectively given in
+pseudo-spin operators $S_{\N\pm}^\x$ are respectively given in
 \eqref{eq:O_N+}, \eqref{eq:O_N-}, and \eqref{eq:S_NX_x}.  Denoting the
-eigenstates of $S_{\N,\pm}^\x$ by
-\begin{align}
-  \ket{\N,\S_\pm} \equiv \f1{\sqrt2} \p{\ket{\N,0} + \ket{\N,\pm}},
-  &&
-  \ket{\N,\A_\pm} \equiv \f1{\sqrt2} \p{\ket{\N,0} - \ket{\N,\pm}},
+eigenstates of $S_{\N\pm}^\x$ by
+\begin{align}
+  \ket{\N,\S_\pm} \equiv \f1{\sqrt2} \p{\ket{\N,0} + \ket{\N\pm}},
+  &&
+  \ket{\N,\A_\pm} \equiv \f1{\sqrt2} \p{\ket{\N,0} - \ket{\N\pm}},
 \end{align}
@@
 \begin{align}
-  \1_{\N,\pm} \equiv \op{\N,0} + \op{\N,\pm}
+  \1_{\N\pm} \equiv \op{\N,0} + \op{\N\pm}
   = \op{\N,\S_\pm} + \op{\N,\A_\pm},
@@
 we can write the state $\rho_N^{(\pm)}\p{t}$ and excited-state
-projectors $\P_{N,\pm}$ in the form
+projectors $\P_{N\pm}$ in the form
 \begin{align}
   \rho_N^{(\pm)}\p{t} = \f1{\Tr\P_{N,0}} \sum_{\abs{\N}=N} \f12
-  \sp{\1_{\N,\pm} + e^{i2t\Omega_{\N,\pm}}\op{\N,\S_\pm}{\N,\A_\pm}
-    + e^{-i2t\Omega_{\N,\pm}}\op{\N,\A_\pm}{\N,\S_\pm}},
+  \sp{\1_{\N\pm} + e^{i2t\Omega_{\N\pm}}\op{\N,\S_\pm}{\N,\A_\pm}
+    + e^{-i2t\Omega_{\N\pm}}\op{\N,\A_\pm}{\N,\S_\pm}},
   \label{eq:rho_N_pm_t}
@@
 \begin{align}
-  \P_{N,\pm} = \sum_{\abs{\N}=N} \f12
-  \sp{\1_{\N,\pm}-\op{\N,\S_\pm}{\N,\A_\pm}-\op{\N,\A_\pm}{\N,\S_\pm}},
+  \P_{N\pm} = \sum_{\abs{\N}=N} \f12
+  \sp{\1_{\N\pm}-\op{\N,\S_\pm}{\N,\A_\pm}-\op{\N,\A_\pm}{\N,\S_\pm}},
 \end{align}
@@
 \begin{align}
-  \bk{\P_{N,\pm}\p{t}}
-  \equiv \Tr\sp{\rho_N^{(\pm)}\p{t}\P_{N,\pm}}
-  = \f12 - \f12 \bk{\cos\p{2t\Omega_{\N,\pm}}}_{\abs{\N}=N},
+  \bk{\P_{N\pm}\p{t}}
+  \equiv \Tr\sp{\rho_N^{(\pm)}\p{t}\P_{N\pm}}
+  = \f12 - \f12 \bk{\cos\p{2t\Omega_{\N\pm}}}_{\abs{\N}=N},
   \label{eq:P_N_pm_t}
@@
   \caption{\footnotesize Net population of the $N$-body orbital
-    excited states $\set{\ket{\N,\pm}}$ after interrogation of an
+    excited states $\set{\ket{\N\pm}}$ after interrogation of an
     initial mixed state $\rho_{N,0}=\P_{N,0}/\Tr\P_{N,0}$ for a
@@
 Figure \ref{fig:time} shows the excited-state population
-$\bk{\P_{N,\pm}\p{t}}$ for several occupation numbers $N$.  With the
+$\bk{\P_{N\pm}\p{t}}$ for several occupation numbers $N$.  With the
 exception of $N=2$, the asymmetric-state populations generally have an
@@
 of many-body state structures, as well as for the preparation of
-long-lived states with multi-partite entanglement
-(i.e.~$\ket{\N,\pm}$) which may be used as a resource to perform
-quantum information processing tasks.
+long-lived states with multi-partite entanglement (i.e.~$\ket{\N\pm}$)
+which may be used as a resource to perform quantum information
+processing tasks.
 
@@
 where $G_{\t{HO}}$ is an effective coupling constant in the harmonic
-trap, $\phi_0^{\t{HO}}\p{x}$ is the corresponding single-particle
+trap, $\phi_0^{\t{HO}}\p{x}$ is the corresponding non-interacting
 ground-state wavefunction, and $\Gamma$ is the gamma function.  The

### 2019-02-11 multi_body_interactions/papers/theory/paper.tex
renaming referee response file, and adding some misc. revisions

@@
 ground state with an incoherent mixture of all nuclear spins, then we
-prepare the mixed state $\rho_{N,0}=\P_{N,0}/\Tr\P_{N,0}$, where
+prepare the mixed state $\rho_{N,0}=\P_{N,0}/\tr\P_{N,0}$, where
 \begin{align}
@@
 \begin{align}
-  \rho_N^{(\pm)}\p{t} = \f1{\Tr\P_{N,0}}
+  \rho_N^{(\pm)}\p{t} = \f1{\tr\P_{N,0}}
   \sum_{\abs{\N}=N} \exp\p{-it\Omega_{\N\pm}S_{\N\pm}^\x}
@@
 \begin{align}
-  \rho_N^{(\pm)}\p{t} = \f1{\Tr\P_{N,0}} \sum_{\abs{\N}=N} \f12
+  \rho_N^{(\pm)}\p{t} = \f1{\tr\P_{N,0}} \sum_{\abs{\N}=N} \f12
   \sp{\1_{\N\pm} + e^{i2t\Omega_{\N\pm}}\op{\N,\S_\pm}{\N,\A_\pm}
@@
   \bk{\P_{N\pm}\p{t}}
-  \equiv \Tr\sp{\rho_N^{(\pm)}\p{t}\P_{N\pm}}
+  \equiv \tr\sp{\rho_N^{(\pm)}\p{t}\P_{N\pm}}
   = \f12 - \f12 \bk{\cos\p{2t\Omega_{\N\pm}}}_{\abs{\N}=N},
@@
 \end{align}
-where $\bk{X}_{\abs{\N}=N}\equiv\sum_{\abs{\N}=N}X/\Tr\P_{N,0}$ is an
+where $\bk{X}_{\abs{\N}=N}\equiv\sum_{\abs{\N}=N}X/\tr\P_{N,0}$ is an
 average of $X$ over all choices of $N$ distinct nuclear spins.
@@
     excited states $\set{\ket{\N\pm}}$ after interrogation of an
-    initial mixed state $\rho_{N,0}=\P_{N,0}/\Tr\P_{N,0}$ for a
+    initial mixed state $\rho_{N,0}=\P_{N,0}/\tr\P_{N,0}$ for a
     reduced time $\tau_N\equiv t\Omega_0/\sqrt{N}$ (i.e.~with real
@@
 population transfer to the excited state
-$\rho_{N,-}\equiv\P_{N,-}/\Tr\P_{N,-}$ for odd $n$, and a collapse
+$\rho_{N,-}\equiv\P_{N,-}/\tr\P_{N,-}$ for odd $n$, and a collapse
 back to the orbital ground state $\rho_{N,0}$ for even $n$, precisely
@@
   also known as the $L_{2,2}$, Hilbert-Schmidt, or Frobenius norm.}
-$\norm{X}_{\t{E}}\equiv\sqrt{\Tr\p{X^\dag X}}$.  This rotation is
+$\norm{X}_{\t{E}}\equiv\sqrt{\tr\p{X^\dag X}}$.  This rotation is
 determined uniquely by enforcing

### 2019-02-11 multi_body_interactions/papers/theory/reviews_NJP/revisions.tex
renaming referee response file, and adding some misc. revisions

@@
   fixed value (as e.g.~in $K^{k\ell}_{mn}$, $K^{\ell,3}_{2,5}$,
-  $\c_{\mu n}$, and $\c_{\mu,\g}$).  To this end, in our last appendix
-  (page 40, line 17) we have replaced all instances of $U_{M,X}$ and
-  $U_{M,X}^{(p)}$ respectively with $U_{MX}$ and $U_{MX}^{(p)}$.
+  $\c_{\mu n}$, and $\c_{\mu,\g}$).  To this end, we have removed the
+  commas on the subscripts ``$M,X$''; ``$N,\pm$''; and ``$\N,\pm$''
+  whenever they appear.
 
@@
 
-  Both examples are correct due to the fact that
+  Both examples are ``correct'' due to the fact that
   $G^{\mu q;\nu r}_{\rho s;\sigma t}$ is self-adjoint, but the latter
-  ordering convention is used throughout the rest of the manuscript.
+  ordering convention is used throughout the rest of the manuscript:
+  $G^{\mu q;\nu r}_{\rho s;\sigma t}$ is gives the coupling constant
+  for destroying two atoms in state $\p{\mu q}+\p{\nu r}$, and
+  creating them in $\p{\rho s}+\p{\sigma t}$.
 
@@
 
+\item For clarity, we have replaced a few instances of the text
+  ``\red{single-particle}'' by ``\green{non-interacting}'' when
+  referring to motional ground states of non-interacting particles;
+  specifically, we have done so on page 9, line 50; on page 13, line
+  5; and on page 35, line 26.
+
 \item Reference 34 of the original manuscript has been published since

### 2019-02-12 multi_body_interactions/papers/theory/reviews_NJP/referee_response_1.tex
updated revisions summary and added definition of 'internal vertex'

--- 2019-02-11 multi_body_interactions/papers/theory/reviews_NJP/referee_1_response.tex
@@
   In section IV (page 18), we have similarly changed the first
-  sentence from (page 18, line line 45):
+  sentence from (page 18, line 45):
 
@@
   have therefore added the a paragraph to section II, following the
-  re-statement of Eq.~(2) in Eq.~(5).  Specifically, the following
-  text has been inserted in between lines 35 and 36 of page 6:
+  re-statement of Eq.~(2) in Eq.~(5).  Specifically, we have inserted
+  the following text between lines 35 and 36 of page 6:
 
@@
     for our eventual development of a low-energy effective theory.}
+
+  Note that the above text is part of a revision made in point
+  \ref{pt:pseudo-potential}, not a ``new'' revision.
 

### 2019-02-12 multi_body_interactions/papers/theory/paper.tex
updated revisions summary and added definition of 'internal vertex'

@@
 interaction terms at order $p$ in $H_{\t{int}}$ can be represented by
-directed graphs containing $p$ internal vertices.  Each internal
-vertex and its associated edges correspond respectively to a coupling
-constant and the associated field operators in $H_{\t{int}}$.  An
-example 2-vertex diagram representing an effective 3-body interaction
-term is provided in figure \ref{fig:diagram}.  All diagrams are read
-from left to right to construct a sequence of operators from right to
-left; the internal vertices of a diagram are thus strictly ordered,
-with the $n$-th internal vertex from the left corresponding to the
-$n$-th interaction Hamiltonian $H_{\t{int}}$ from the right in
-\eqref{eq:H_int_1_2} or \eqref{eq:H_int_3}.  Solid (dashed) lines
-represent field operators acting on the lowest (arbitrary) motional
-states.  Spatial overlap factors at each vertex are determined by the
-motional states of the edges which connect to (i.e.~field operators
-associated with) that vertex.  While it is possible to construct
-explicit rules for determining the energetic suppression factors
-(i.e.~from $\I$) of the term represented by a diagram, these factors
-are most easily determined by examination of the effective
-Hamiltonians in \eqref{eq:H_int_1_2} and \eqref{eq:H_int_3}.
+directed graphs containing $p$ vertices with degree greater than one,
+which we call {\it internal vervices}.  Each internal vertex and its
+associated edges correspond respectively to a coupling constant and
+the associated field operators in $H_{\t{int}}$.  An example 2-vertex
+diagram representing an effective 3-body interaction term is provided
+in figure \ref{fig:diagram}.  All diagrams are read from left to right
+to construct a sequence of operators from right to left; the internal
+vertices of a diagram are thus strictly ordered, with the $n$-th
+internal vertex from the left corresponding to the $n$-th interaction
+Hamiltonian $H_{\t{int}}$ from the right in \eqref{eq:H_int_1_2} or
+\eqref{eq:H_int_3}.  Solid (dashed) lines represent field operators
+acting on the lowest (arbitrary) motional states.  Spatial overlap
+factors at each vertex are determined by the motional states of the
+edges which connect to (i.e.~field operators associated with) that
+vertex.  While it is possible to construct explicit rules for
+determining the energetic suppression factors (i.e.~from $\I$) of the
+term represented by a diagram, these factors are most easily
+determined by examination of the effective Hamiltonians in
+\eqref{eq:H_int_1_2} and \eqref{eq:H_int_3}.
 

### 2019-02-13 multi_body_interactions/papers/theory/reviews_NJP/revision_summary.tex
renamed revision summary for multi-body NJP article

\documentclass[preprint]{revtex4-1}

% linking references
\usepackage{hyperref}
\hypersetup{
  breaklinks=true,
  colorlinks=true,
  linkcolor=blue,
  urlcolor=cyan,
}

% general physics / math packages and commands
\usepackage{physics,amsmath,amssymb,braket,dsfont}
\renewcommand{\t}{\text} % text in math mode
\newcommand{\f}{\dfrac} % shorthand for fractions
\newcommand{\p}[1]{\left(#1\right)} % parenthesis
\renewcommand{\sp}[1]{\left[#1\right]} % square parenthesis
\renewcommand{\set}[1]{\left\{#1\right\}} % curly parenthesis
\newcommand{\bk}{\braket} % shorthand for braket

\renewcommand{\d}{\text{d}}
\newcommand{\g}{\text{g}}
\newcommand{\e}{\text{e}}
\newcommand{\x}{\text{x}}
\newcommand{\y}{\text{y}}
\newcommand{\z}{\text{z}}

\renewcommand{\c}{\hat{c}}
\newcommand{\n}{\hat{n}}

\newcommand{\A}{\mathcal{A}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\G}{\mathcal{G}}
\renewcommand{\H}{\mathcal{H}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\K}{\mathcal{K}}
\renewcommand{\L}{\mathcal{L}}
\newcommand{\M}{\mathcal{M}}
\newcommand{\N}{\mathcal{N}}
\renewcommand{\O}{\mathcal{O}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\Q}{\mathcal{Q}}
\renewcommand{\S}{\mathcal{S}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\1}{\mathds{1}}

% inline lists
\usepackage[inline]{enumitem}
\setlist[enumerate]{leftmargin=*} % nice margins for enumerate

% for feynman diagrams
\usepackage{tikz,tikz-feynman}
\tikzset{
  baseline = (current bounding box.center)
}
\tikzfeynmanset{
  compat = 1.1.0,
  every feynman = {/tikzfeynman/small}
}
\newcommand{\shrink}[1]{\scalebox{0.8}{#1}} % for smaller diagrams

% color definitions (used in a figure)
\usepackage{xcolor}
\newcommand{\blue}[1]{\textcolor{blue}{#1}}
\newcommand{\red}[1]{\textcolor{red}{#1}}
\newcommand{\green}[1]{\textcolor{green}{#1}}
\definecolor{lightblue}{RGB}{31,119,180}
\definecolor{orange}{RGB}{255,127,14}
\definecolor{green}{RGB}{44,160,44}
\definecolor{lightred}{RGB}{214,39,40}

% for strikeout text
% normalem included to prevent underlining titles in the bibliography
\usepackage[normalem]{ulem}

% for numbering points in response to referees
\newcounter{point}
\newcommand{\step}{\stepcounter{point}\setcounter{enumi}{0}}

\begin{document}

Dear editors,

In this document, we summarize all revisions made to our manuscript
since its initial submission.  Revisions made in our responses to
Referee 1 (Referee 2) are numbered R1.$X$.$Y$ (R2.$X$.$Y$),
corresponding to the $Y$-th revision in point $X$ of our response.  In
addition, we summarize a few minor revisions made to refine wording or
fix errors in the original manuscript; these revisions are numbered
R3.$Z$.

Excerpts from our original manuscript are written in \red{red}, and
excerpts from the revised version of our manuscript is written in
\green{green}.  All page, line, equation, and reference numbers
generally refer to those in our original manuscript, unless explicitly
stated otherwise.

\section*{Revisions in response to Referee 1}

\setcounter{point}{1}
\begin{enumerate}[label=(R1.\arabic{point}.\arabic*)]

  \step
\item We have inserted a sentence on page 1, line 34:

  \red{... nuclear spin levels.  We fully characterize...}

  which now reads:

  \green{... nuclear spin levels.  Our theory is limited to the
    experimental regime of (i) a deep lattice, with (ii) at most one
    atom occupying each nuclear spin state on any lattice site.  The
    latter restriction is a consequence of initial ground-state
    preparation.  We fully characterize...}

\item The following text on page 4, line 22:

  \red{Specifically, we consider the preparation of isolated few-body
    systems in the deep-lattice limit, ...}

  now reads:

  \green{Specifically, we consider ground-state preparation of
    isolated few-body systems in the deep-lattice limit, ...}

\item We have appended a sentence to the text on page 4, line 31:

  \red{... through the exquisite capabilities with OLCs.}

  which now reads:

  \green{... through the exquisite capabilities with OLCs.  Our theory
    is limited to the experimental regime of at most one atom
    occupying each nuclear spin state on any lattice site, which is a
    consequence of the experimental protocol which starts with all
    atoms in the ground state.}

\item We have rewritten a paragraph on page 4, line 33:

  \red{Though effective multi-body interactions have previously been
    studied in the context of harmonically [35, 36] and
    lattice-confined [37] neutral bosons prepared in a single
    hyperfine state, our work deals for the first time with fermions
    that have internal degrees of freedom and multiple collisional
    parameters.  Nonetheless, due to the SU($N$) symmetry of these
    collisions, we are able to find a simple way to express effective
    multi-body Hamiltonians and fully characterize their low-lying
    many-body energy eigenstates.  While some past work has detected
    experimental signatures of multi-body interactions in the form of
    quantum phase revivals [38], we perform a direct comparison of the
    many-body interaction energies predicted by our low-energy
    effective theory with experimental measurements of
    density-dependent atomic energy spectra performed in ref.~[34],
    similarly to the measurements with bosons performed in ref.~[39].}

  which now reads:

  \green{Though effective multi-body interactions have previously been
    studied in the context of harmonically [35, 36] and
    lattice-confined [37] neutral bosons prepared in a single
    hyperfine state, our work deals for the first time with fermions
    that have internal degrees of freedom and multiple collisional
    parameters.  Some past work has detected experimental signatures
    of multi-body interactions in the form of quantum phase revivals
    [39].  We instead compare the many-body interaction energies
    predicted by our low-energy effective theory to the experimental
    measurements of the density-dependent orbital excitation spectra
    performed in ref.~[34], similarly to the measurements with bosons
    performed in ref.~[38].  To facilitate this comparison of
    excitation spectra and to characterize the low-lying excitations
    in our effective theory, we consider a restriction of our theory
    to states with at most one orbital excitation per lattice site.
    In this case, we find that the SU($N$) symmetry of atomic
    collisions allow the effective multi-body interactions to take a
    remarkably simple form.}

\item We have revised Eq.~(9) and added some discussion, replacing
  (page 7, line 48):

  \red{Our low-energy effective theory exhibits SU($N$)-symmetric
    multi-body interactions, such that the effective interaction
    Hamiltonian can be written in the form
    \begin{align*}
      H_{\t{int}}^{\t{eff}} = \sum_{M=2}^{2I+1} H_M, \tag{9}
    \end{align*}
    where $H_M$ is an $M$-body Hamiltonian, $I$ is the total nuclear
    spin of each atom (e.g.~$I=9/2$ for ${}^{87}$Sr), and the sum
    terminates at $2I+1$ because this is the largest number of atoms
    which may initially occupy a single lattice site.  Focusing on the
    low-lying orbital state excitations of this theory, we find that
    SU($N$) symmetry allows us to express multi-body Hamiltonians in
    the simple form}

  by:

  \green{Our low-energy effective theory exhibits SU($N$)-symmetric
    multi-body interactions, such that the effective interaction
    Hamiltonian can be written in the form
    \begin{align*}
      H_{\t{int}}^{\t{eff}} = \sum_{M=2}^{2I+1} \sum_{p\ge1} H_M^{(p)},
      \tag{9}
    \end{align*}
    where $H_M^{(p)}$ is an $M$-body Hamiltonian of order $p$ in the
    coupling constants $G_X$, and $I$ is the total nuclear spin of
    each atom (e.g.~$I=9/2$ for ${}^{87}$Sr).  The sum terminates at
    $2I+1$ because this is the largest number of atoms which may
    initially occupy a single lattice site.  We explicitly compute all
    of the $M$-body Hamiltonians $H_M\equiv\sum_p H_M^{(p)}$ through
    order $p=3$, yielding effective two-, three-, and four-body
    interactions.  To (i) facilitate a comparison with the
    experimental measurements of many-body orbital excitation spectra
    performed in ref.~[34] and (ii) characterize the low-lying
    excitations in our effective theory, we additionally restrict the
    multi-body Hamiltonians $H_M$ to states with at most one orbital
    excitation per lattice site.  Under this restriction, we find that
    the SU($N$) symmetry of atomic collisions allows us to express all
    multi-body Hamiltonians in the simple form}

\item The text (page 18, line 45):

  \red{Current experiments with ultracold ${}^{87}$Sr on a lattice can
    coherently address ground states and their low-lying orbital
    excitations for up to five atoms per lattice site [34].  Due to
    the SU($N$) symmetry of inter-atomic interactions, manifest in the
    fact that all coupling constants are independent of nuclear spin,
    we can write all effective $M$-body Hamiltonians which address
    these states in the form}

  now reads:

  \green{Current experiments with ultracold ${}^{87}$Sr on a lattice
    can coherently address ground states and single orbital
    excitations of up to five atoms per lattice site [34].  Due to the
    SU($N$) symmetry of inter-atomic interactions, manifest in the
    fact that all coupling constants are independent of nuclear spin,
    a restriction of the $M$-body Hamiltonians $H_M=\sum_p H_M^{(p)}$
    to the subspace of experimentally addressed states takes the form}

\item The text (page 28, line 40):

  \red{Working in the deep-lattice limit, we have derived a low-energy
    effective theory of such atoms.  This theory exhibits emergent
    multi-body interactions that inherit the SU($N$) symmetry of the
    atoms' bare, two-body interactions.  The SU($N$) symmetry of
    $M$-body Hamiltonians in the effective theory has allowed us to
    express them in a simple form, and fully characterize their
    corresponding eigenstates and spectra.}

  now reads:

  \green{Working in the deep-lattice limit and the experimental regime
    of at most one atom occupying each nuclear spin state on any
    lattice site, we have derived a low-energy effective theory of
    these atoms.  Our theory exhibits emergent multi-body interactions
    that inherit the SU($N$) symmetry of the bare two-body
    interactions.  Considering a restriction of our theory to the
    subspace of at most one orbital excitation per lattice site, we
    found that the SU($N$) symmetry of all $M$-body Hamiltonians
    allowed us to express them in a simple form, and to fully
    characterize their eigenstates and spectra.}

\item We have added a considerable amount of text to the beginning of
  section II, such that the opening (page 5, line 19):

  \red{Consider a collection of fermionic atoms with total nuclear
    spin $I$ and two atomic orbital states loaded into a
    translationally invariant 3-D optical lattice.  While an external
    trapping potential will generally break discrete translational
    symmetry of the lattice, any background inhomogeneity can be made
    negligible by spectroscopically addressing a sufficiently small
    region of the lattice [34].  Throughout this paper, we work
    strictly in the deep-lattice regime with negligible tunneling
    between lattice sites.  We also assume a so-called
    ``magic-wavelength'' lattice for which both atomic orbital states
    experience identical lattice potentials [40].}

  now reads:

  \green{The work in this paper is closely tied to the experimental
    work reported in ref.~[34]; we begin with a short summary of the
    relevant experimental procedures therein.  The experiment begins
    by preparing a degenerate gas of $10^4$-$10^5$ (fermionic)
    ${}^{87}$Sr atoms in a uniform mixture of their ten nuclear spin
    states and at $\sim0.1$ of their fermi temperature ($\sim10$
    nanokelvin) [8, 40].  This gas is loaded into a primitive cubic
    optical lattice at the ``magic wavelength'' for which both ground
    (${}^1S_0$) and first-excited (${}^3P_0$) electronic orbital
    states of the atoms experience the same lattice potential [41].
    Lattice depths along the principal axes of the lattice are roughly
    equal in magnitude, with a geometric mean that can be varied from
    30 to 80 $E_{\t{R}}$, where
    $E_{\t{R}}\approx3.5\times2\pi~\t{kHz}$ is the lattice photon
    recoil energy of the atoms (with the reduced Planck constant
    $\hbar=1$ throughout this paper).  These lattice depths are
    sufficiently large as to neglect tunneling on the time scales
    relevant to the experiment.  The temperature of the atoms is also
    low enough to neglect thermal occupation of motional states
    outside the ground-state manifold.}

  \green{Once loaded into an optical lattice, atoms are addressed by
    an external (``clock'') interrogation laser with an ultranarrow
    (26 mHz) linewidth, detuned by $\Delta$ from the single-atom
    ${}^1S_0-{}^3P_0$ transition frequency $\omega_0$.  After a fixed
    interrogation time, the experiment turns off the interrogation
    laser, removes all ground-state (${}^1S_0$) atoms from the
    lattice, and uses absorption imaging to count the remaining
    excited-state (${}^3P_0$) atoms.  Non-interacting atoms in
    singly-occupied lattice sites feature the typical single particle
    lineshape peaked at $\Delta=0$.  The lineshapes of
    multiply-occupied lattice sites, meanwhile, are shifted by
    inter-atomic interactions, which results in spectroscopic peaks
    (i.e.~local maxima in excited-state atom counts) away from
    $\Delta=0$.  A sweep across different detunings $\Delta$ (on the
    scale of inter-atomic interaction energies) thus constitutes a
    measurement of the many-body orbital excitation spectrum.  We note
    that this spectroscopic protocol addresses only singly-excited
    orbital states of lattice sites.  Doubly-excited states are
    off-resonant due to (i) the interaction-induced non-linearity
    ($\sim$kHz) of orbital excitation energies, and (ii) the
    ultranarrow linewidth ($\sim$mHz) of the interrogation laser.}

  \green{While an external trapping potential will generally break
    discrete translational symmetry of the lattice, any background
    inhomogeneity can be made negligible by spectroscopically
    addressing a sufficiently small region of the lattice [34].
    Throughout this paper, we work strictly in the deep-lattice regime
    with negligible tunneling between lattice sites.  We also neglect
    any lattice inhomogeneities and assume that both atomic orbital
    states (i.e.~${}^1S_0$ and ${}^3P_0$) experience identical lattice
    potentials.}

\item To reflect the above change, the text (page 4, line 52):

  \red{In section II we provide an overview of the relevant one- and
    two-body physics of ultracold atoms on a lattice, and summarize
    our main result of effective multi-body interactions between these
    atom.}

  now reads:

  \green{In section II we summarize the experimental procedures
    relevant to our work, provide an overview of the one- and two-body
    physics of ultracold atoms in a deep lattice, and preview our main
    technical results.}

\item We have removed the (now redundant) parenthetical text on page
  5, line 45:

  \red{(with the reduced Planck constant $\hbar=1$ throughout this
    paper)}

  \step
\item We have re-written the paragraph on page 7, lines 20--47:

  \red{We are interested in the ground and low-lying orbital
    eigenstates of several interacting atoms on a single lattice site,
    as such states can be readily prepared and coherently addressed in
    current experiments with ultracold atoms [34].  Although these
    experiments operate at temperatures sufficiently low for
    negligible thermal occupation of excited motional states, simply
    neglecting excited single-particle motional states fails to
    reproduce the observed spectra of multiply-occupied lattice sites.
    The reason for this failure lies in the fact that the bare
    two-body Hamiltonian $H_{\t{int}}$ in (7) is not diagonal in the
    basis of single-particle motional states.  Even at zero
    temperature, therefore, the presence of excited motional states
    modifies many-body atomic energy spectra when the magnitude of
    off-diagonal elements of $H_{\t{int}}$ approach the scale of
    low-lying single-particle motional state energy gaps.  To address
    this problem, we develop a low-energy effective theory for
    ultracold two-level fermionic atoms on a lattice.  As our work is
    motivated by progress in experiments which initialize all atoms in
    identical (i.e.~ground) orbital and motional states, to simplify
    our theory we assume that any $N$ atoms on a single lattice site
    occupy $N$ nuclear spin states.  Multiple occupation of a single
    nuclear spin state on the same lattice site is initially forbidden
    by fermionic statistics, and subsequently cannot be violated in
    the absence of tunneling and hyperfine coupling.}

  In place of the above paragraph, we now have the text:

  \green{Current experiments with ${}^{87}$Sr can prepare up to five
    atoms in the same (ground) orbital state on a single lattice site,
    and coherently address states with a single orbital excitation per
    lattice site [34].  At ultracold temperatures well below the
    single-atom motional excitation energies $\Delta_n\equiv E_n-E_0$
    for $n>0$, atoms only occupy their motional ground state in the
    lattice.  For this reason, it is common to map the description of
    these atoms onto a single-band Hubbard model that captures all
    dynamics within the subspace of motional ground states of
    non-interacting atoms, i.e.~with single-atom wavefunctions
    $\phi_{i,0}$.  Interactions, however, modify atoms' motional
    ground-state wavefuctions.  The true motional ground state of a
    collection of interacting atoms is then an admixture of the
    non-interacting motional eigenstates, and a naive Hubbard model
    that assumes atomic wavefunctions $\phi_{i,0}$ will fail to
    reproduce the interacting atoms' orbital excitation spectrum.
    Formally, corrections to the spectrum of interacting atoms can be
    accounted for by a perturbative treatment of far-off-resonant
    terms in the interaction Hamiltonian of (7) that create atoms in
    excited motional states,
    e.g.~$\sim\c_{n\mu s}^\dag \c_{0,\nu t}^\dag \c_{0,\nu r}
    \c_{0,\mu q}$ with $n>0$.  These corrections can be understood
    through interaction-induced {\it virtual} occupation of higher
    bands (i.e.~excited motional states), which becomes relevant as
    more atoms occupy the same lattice site, such that their
    interaction energy becomes non-negligible compared to the motional
    excitation energies $\Delta_n$.}

  \green{In order to recast interaction-induced modifications to
    orbital excitation spectra as corrections to the simple Hubbard
    model (i.e.~computed using the non-interacting ground-state
    wavefunctions $\phi_{i,0}$), we develop a low-energy effective
    theory of interacting AEAs in a deep lattice.  To simplify our
    theory, we assume that any $N$ atoms on a single lattice site
    occupy distinct nuclear spin states.  This assumption applies for
    any experimental protocol in which all atoms are initially
    prepared in their orbital and motional ground states (as e.g.~in
    ref.~[34]).  In this case, multiple occupation of a single nuclear
    spin state on any given lattice site is initially forbidden by
    fermionic statistics.  Subsequent violation of this condition
    cannot occur in the absence of inter-site effects or hyperfine
    coupling between nuclear spin states, as is the case in the
    experiment in ref.~[34].}

\item We have changed the text (page 9, line 45):

  \red{The naive idea of neglecting all excited states and using
    $H_{\t{int}}$ directly to describe the spectrum of single-particle
    motional ground states is thus equivalent to truncating our
    expansion for $H_{\t{int}}^{\t{eff}}$ at first order.}

  to:

  \green{Writing down a single-band Hubbard model that simply neglects
    excited atomic motional states and uses $H_{\t{int}}$ directly to
    describe the orbital spectrum of interacting atoms is thus
    equivalent to truncating our expansion for $H_{\t{int}}^{\t{eff}}$
    at first order.}

  \step
\item Eq.~(10) and the following discussion (page 8, line 10):

  \red{\begin{align*} H_M = \sum_{\abs{\set{\mu_j}}=M} \p{U_{M,\g}
        \prod_{\alpha=1}^M \n_{\mu_\alpha,\g} + U_{M,+} \n_{\mu_1,\e}
        \prod_{\alpha=2}^M \n_{\mu_\alpha,\g} + U_{M,-}
        \c_{\mu_1,\g}^\dag \c_{\mu_2,\e}^\dag \c_{\mu_2,\g}
        \c_{\mu_1,\e} \prod_{\alpha=3}^M \n_{\mu_\alpha,\g}},
      \tag{10}
    \end{align*}
    where $\c_{\mu s}\equiv\c_{0,\mu s}$ is a fermionic annihilation
    operator for an atom in the motional ground state with nuclear
    spin $\mu$ and orbital state $s$;
    $\n_{\mu s}\equiv \c_{\mu s}^\dag \c_{\mu s}$ is a number
    operator; and the sum is performed over all choices of nuclear
    spins $\mu_j$ with $j=1,2,\cdots,M$ for which all $\mu_j$ are
    distinct, or equivalently all choices of $\mu_j$ for which the set
    $\set{\mu_j}$ contains $M$ elements, for a total of
    ${2I+1\choose M}\times\p{M!}$ nuclear spin combinations.  The
    coefficients $U_X$ can be expressed in terms of the coupling
    constants of the bare two-body interaction Hamiltonian (see
    Appendix F).  The key feature of the $M$-body interactions in (10)
    is that they take a form identical to the two-body interactions
    between motional ground-state atoms, but with the addition of
    $M-2$ spectator atoms.}

  now reads:

  \green{\begin{align*}
      H_M = \sum_{\abs{\set{\mu_j}}=M}
      H_2^{(\mu_1,\mu_2)} \prod_{\alpha=3}^M \n_{\mu_\alpha,\g},
      \tag{10}
    \end{align*}
    where $H_2^{(\mu_1,\mu_2)}$ is a two-body Hamiltonian addressing
    atoms with nuclear spin $\mu_1,\mu_2$; and
    $\n_{\mu s}=\c_{\mu s}^\dag\c_{\mu s}$ is a number operator for
    atoms with nuclear state $\mu$ and orbital state $s$.  The sum in
    (10) is performed over all choices of nuclear spins $\mu_j$ with
    $j=1,2,\cdots,M$ for which all $\mu_j$ are distinct, or
    equivalently all choices of $\mu_j$ for which the set
    $\set{\mu_j}$ contains $M$ elements, for a total of
    ${2I+1\choose M}\times\p{M!}$ nuclear spin combinations.  The key
    feature of the $M$-body interactions in (10) is that they
    ultimately take the same form as two-body interactions, but with
    the addition of $M-2$ spectator atoms.}

\item Similarly, Eq.~(42) has been changed from (page 18, line 54):

  \red{\begin{align*}
      H_M = \sum_{\abs{\set{\mu_j}}=M}
      \p{U_{M,\g} \prod_{\alpha=1}^M \n_{\mu_\alpha,\g}
        + U_{M,+} \n_{\mu_1,\e} \prod_{\alpha=2}^M \n_{\mu_\alpha,\g}
        + U_{M,-} \c_{\mu_1,\g}^\dag \c_{\mu_2,\e}^\dag
        \c_{\mu_2,\g} \c_{\mu_1,\e} \prod_{\alpha=3}^M \n_{\mu_\alpha,\g}},
      \tag{42}
    \end{align*}}

  to:

  \green{\begin{align*}
      H_M = \sum_{\abs{\set{\mu_j}}=M}
      \p{U_{M,\g} \n_{\mu_1,\g} \n_{\mu_2,\g}
        + U_{M,+} \n_{\mu_1,\e} \n_{\mu_2,\g}
        + U_{M,-} \c_{\mu_1,\g}^\dag \c_{\mu_2,\e}^\dag
        \c_{\mu_2,\g} \c_{\mu_1,\e}}
      \prod_{\alpha=3}^M \n_{\mu_\alpha,\g},
      \tag{44}
    \end{align*}}

  \step
\item We have inserted the following text between lines 35 and 36 of
  page 6:

  \green{Note that the Hamiltonian in (5) is not the true microscopic
    interaction Hamiltonian of AEAs, but rather a generic form for a
    low-energy effective field theoretic description of two-body
    interactions [17, 21, 35--37, 42--44].  There are therefore two
    important points to keep in mind concerning our use of (5) to
    describe two-body interactions.  First, the use of effective field
    theory generically gives rise to divergences that must be dealt
    with either through regularization, e.g.~of the zero-range
    interaction potential implicitly assumed in the expression of (5)
    [45], or through renormalization of the coupling constants in the
    theory.  We chose the latter approach, as we will in any case find
    it convenient to renormalize the coupling constants in the
    effective theory developed in section III.  The choice of method
    to regulate divergences has no effect on the underlying physics.}

  \green{Second, (5) is only the first term in a low-energy expansion
    of two-body interactions in effective field theory, which
    generally includes additional terms containing derivatives of
    field operators.  Derivative terms correspond to the dependence of
    two-body scattering on the relative momentum $k$ of particles
    involved, with $k\to0$ in the zero-energy limit.  In the present
    case of $s$-wave scattering, the leading dependence of the
    two-body interaction Hamiltonian on the relative momentum $k$ can
    be captured by use of an energy-dependent pseudo-potential, which
    amounts to using a $k$-dependent effective scattering length [46].
    This effective scattering length can be determined by expanding
    the $s$-wave collisional phase shift in powers of the relative
    momentum $k$ [45, 47].  Details of this expansion will depend on
    the characteristic length scale of finite-range interactions.  In
    our work, these corrections to (5) will be relevant only for the
    calculation of two-body interaction energies, appearing at third
    order in the coupling constants $G_X$.  As we are primarily
    interested in $M$-body interactions for $M\ge3$, we defer this
    calculation to Appendix D.  We note that our approach of using an
    unregularized contact potential, renormalizing coupling constants,
    and separately accounting for momentum-dependent scattering is
    essentially the same as the approach used for similar calculations
    in refs.~[35--37].  While this approach does not provide insight
    into the microscopic nature of atomic interactions, it is suitable
    for the phenomenological description of these interactions, and
    particular for our eventual development of a low-energy effective
    theory.}

\item The text (page 12, line 15):

  \red{In general, the sums over excited states in the loop diagrams
    of (15) will diverge [35], which implies the need to renormalize
    the coupling constants in our theory.  Even if these diagrams do
    not diverge, we would like to express the effective two-body
    interactions in terms of net, physically observable two-body
    coupling constants as on the right-hand side of (15), rather than
    in terms of long sums at all orders of the bare coupling
    constants.  We therefore renormalize our coupling constants by
    introducing counter-terms $\tilde G_X$ into the effective
    interaction Hamiltonians via $G_X\to G_X+\tilde G_X$.  The values
    of these counter-terms are fixed by enforcing that $G_X$ is equal
    to the net two-body interaction strength.}

  now reads:

  \green{On physical grounds, the net two-body interaction must
    clearly be finite, but individual sums over the excited states in
    loop diagrams of (15) may generally diverge [35].  These
    divergences ultimately appear due to our use of effective field
    theory to describe inter-atomic interactions in (2), (5), and (7),
    rather than a detailed microscopic description of two-atom
    scattering.  Divergences of this sort are a generic feature of
    field theories, and can be dealt with using standard techniques
    such as renormalization.  We therefore renormalize our coupling
    constants by introducing counter-terms $\tilde G_X$ into the
    interaction Hamiltonian.}

\item The text (page 13, line 24):

  \red{Second, the overview we provided in section II, and
    consequently our result in (17), neglects the effect of
    momentum-dependent two-body scattering. When the effective range
    of momentum-dependent interactions is comparable to the scattering
    lengths $a_X$, as is the case for ultracold ${}^{87}$Sr, these
    interactions are third order in the coupling constants $G_X$. Just
    as the $\O\p{G^2}$ counter-terms do not affect $M$-body
    interactions until next-to-leading order (NLO), the $\O\p{G^3}$
    momentum-dependent interactions do not come into play until next-
    next-leading order (NNLO).}

  now reads:

  \green{Finally, our result in (17) neglects the effect of
    momentum-dependent two-body scattering.  When the effective range
    of interactions is comparable to the scattering lengths $a_X$, as
    is the case for ultracold ${}^{87}$Sr, these momentum-dependent
    effects are third order in the coupling constants $G_X$.  Just as
    the $\O\p{G^2}$ counter-terms do not affect $M$-body interactions
    until next-to-leading order (NLO), the $\O\p{G^3}$
    momentum-dependent terms do not come into play until
    next-next-leading order (NNLO).}

  \step
\item We have replaced the following two paragraphs (page 8, line 36):

  \red{As the bare two-body Hamiltonian $H_{\t{int}}$ is not diagonal
    with respect to single-particle motional states, the problem of
    determining many-body spectra and eigenstates nominally involves
    all motional degrees of freedom of the atoms involved.  If we are
    interested in addressing internal atomic states at zero
    temperature, however, in principle we need only to consider the
    motional ground state associated with each many-body internal
    atomic state.  We can then ignore all excited motional states, as
    they will be neither thermally occupied nor externally
    interrogated.  This procedure makes the many-body problem more
    tractable by greatly reducing the dimensionality of the relevant
    Hilbert space.}

  \red{We denote the space of all internal states of atoms in
    single-particle, non-interacting (multi-particle, interacting)
    motional ground states by $\H_{\t{ground}}^{\t{single}}$
    ($\H_{\t{ground}}^{\t{multi}}$).  For sufficiently weak
    interactions, one can construct a unitary transformation $U$
    acting on the full many-body Hilbert space which rotates
    $\H_{\t{ground}}^{\t{multi}}$ into $\H_{\t{ground}}^{\t{single}}$
    [43].  The unitary $U$ can then be used to construct an {\it
      effective Hamiltonian} $H_{\t{int}}^{\t{eff}}$ on
    $\H_{\t{ground}}^{\t{single}}$ that reproduces the spectrum of
    $H_{\t{int}}$ on $\H_{\t{ground}}^{\t{multi}}$, namely
    $H_{\t{int}}^{\t{eff}}=UH_{\t{int}}U^\dag$.  This method for
    constructing an effective theory is commonly known as the
    Schrieffer-Wolff transformation, named after the authors of its
    celebrated application in relating the Anderson and Kondo models
    of magnetic impurities in metals [44].}

  by:

  \green{The net Hamiltonian $H = H_0 + H_{\t{int}}$ for interacting
    AEAs on a lattice is not diagonal with respect to single-particle
    motional state indices (e.g.~$n\in\mathbb{N}_0^3$).  The problem
    of determining interacting atoms' orbital excitation spectrum
    therefore nominally involves all atomic motional degrees of
    freedom.  At zero temperature, however, each orbital state of a
    collection of interacting atoms is associated with a single
    motional ground state.  In order to compute an orbital excitation
    spectrum at zero temperature, in principle we need to identify
    this motional ground state.  We can then ignore all excited
    motional states, which will be neither thermally occupied nor
    externally interrogated.  Such a procedure would drastically
    reduce the dimensionality of the Hilbert space necessary to
    describe the atoms, thereby greatly simplifying any description of
    the atoms' orbital spectrum and internal (i.e.~nuclear and
    orbital) dynamics.  In practice, however, identifying the motional
    ground states of interacting atoms and writing down a Hamiltonian
    restricted to this subspace is a very difficult process to carry
    out analytically.}

  \green{We denote the motional ground-state subspace of the
    non-interacting Hamiltonian $H_0$ by
    $\H_{\t{ground}}^{\t{single}}$, and the motional ground-state
    subspace of the interacting Hamiltonian $H = H_0 + H_{\t{int}}$ by
    $\H_{\t{ground}}^{\t{multi}}$.  That is, all atomic wavefunctions
    for states within $\H_{\t{ground}}^{\t{single}}$ are described by
    $\phi_{i,0}$, while the atomic wavefunctions for states within
    $\H_{\t{ground}}^{\t{multi}}$ are generally unknown, and are in
    principle determined by minimizing the energy of a state with
    respect to its motional degrees of freedom.  Both
    $\H_{\t{ground}}^{\t{single}}$ and $\H_{\t{ground}}^{\t{multi}}$
    are subspaces of the full Hilbert space $\H_{\t{full}}$.  When
    interactions are sufficiently weak compared to the spectral gap
    $\Delta$ between $\H_{\t{ground}}^{\t{single}}$ and its orthogonal
    complement $\H_{\t{full}}\setminus\H_{\t{ground}}^{\t{single}}$,
    one can identify a particular unitary operator $U$ (acting on the
    full Hilbert space $\H_{\t{full}}$) which rotates
    $\H_{\t{ground}}^{\t{multi}}$ into $\H_{\t{ground}}^{\t{single}}$
    [49].  This unitary $U$ can be used to construct an {\it effective
      Hamiltonian} $H_{\t{eff}} = U H U^\dag$ with two key properties:
    (i) $H_{\t{eff}}$ is diagonal in the same (known) basis as the
    non-interacting Hamiltonian $H_0$, and (ii) the spectrum of
    $H_{\t{eff}}$ on $\H_{\t{ground}}^{\t{single}}$ is identical to
    that of the interacting Hamiltonian $H$ on
    $\H_{\t{ground}}^{\t{multi}}$.  The use of an effective
    Hamiltonian $H_{\t{eff}}$ thus overcomes the need to identify
    $\H_{\t{ground}}^{\t{multi}}$ in order to compute the orbital
    spectrum of $H$ at zero temperature.  This method for constructing
    an effective theory is commonly known as the Schrieffer-Wolff
    transformation, named after the authors of its celebrated
    application in relating the Anderson and Kondo models of magnetic
    impurities in metals [50].}

\item We have replaced the text (page 10, line 19):

  \red{Finally, the effective theory involves no corrections to the
    many-body energy eigenstates.  Including such corrections would
    invalidate the effective theory, which preserves the eigenstates
    of the non-interacting Hamiltonian $H_0$ on
    $\H_{\t{ground}}^{\t{single}}$.}

  by:

  \green{Finally, the effective theory involves no corrections to the
    non-interacting many-body energy eigenstates; the purpose of
    constructing the effective Hamiltonian
    $H_{\t{eff}} = H_0 + H_{\t{int}}^{\t{eff}}$ is to reproduce, on
    the known eigenstates of the non-interacting Hamiltonian $H_0$
    within $\H_{\t{ground}}^{\t{single}}$, the spectrum of the
    interacting Hamiltonian $H = H_0 + H_{\t{int}}$ on
    $\H_{\t{ground}}^{\t{multi}}$.  ``Correcting'' the eigenstates of
    the non-interacting Hamiltonian $H_0$ on
    $\H_{\t{ground}}^{\t{single}}$ thus invalidates the effective
    theory.}

  \step
  \step
\item We have appended the following text to section III D, after
  Eq.~(32) on page 16:

  \green{Note that although (22) is a loop diagram, the associated
    factors $\alpha_3^{(3)}$ and $\alpha_5^{(3)}$ in $H_3^{(3)}$ are
    finite.  At large $n$, atoms become free particles for which $n$
    essentially indexes discrete momentum states.  These atoms thus
    have an energy which asymptotically scales as
    $E_n\sim n^2\equiv n_\x^2+n_\y^2+n_\z^2$.  Furthermore, the
    oscillatory behavior of atomic wavefunctions with increasing
    motional state indices $\ell,m$ implies that the overlap integral
    $K_{\ell m}$ becomes sharply peaked at $\ell\approx m$ as $\ell$
    and $m$ get large.  The asymptotic behavior of $\alpha_3^{(3)}$ at
    large $\ell,m,n$ is therefore
    \begin{align*}
      \alpha_3^{(3)}
      \sim \int \f{\d^3\ell~\d^3m~\d^3n}{\p{\ell^2+m^2}\p{\ell^2+n^2}}
      ~\delta\p{\ell-m}\delta\p{\ell-n}
      \sim \int \f{\d^3\ell}{\ell^4}
      \sim \int_{\ell_{\t{min}}}^\infty \f{\d\ell}{\ell^2}
      \sim \f1{\ell_{\t{min}}},
      \tag{33}
    \end{align*}
    where in the last integral we changed to spherical coordinates,
    and $\ell_{\t{min}}^2$ is the minimum value of $\ell^2$ for which
    (i) the energy $E_\ell\sim\ell^2$, and (ii) the integral
    expression in (33) is a good approximation to the corresponding
    sum in (30).  Note that the introduction of $\ell_{\t{min}}$
    amounts to neglecting a finite number of terms in the sum over
    $\ell,m,n$ in (30), whose contribution to the value of
    $\alpha_3^{(3)}$ is finite.  Convergence of $\alpha_5^{(3)}$ is
    similarly guaranteed by the fact that the overlap integral $K_n$
    does not asymptotically grow with increasing $n$, such that
    $\alpha_5^{(3)}$ asymptotically behaves as
    \begin{align*}
      \alpha_5^{(3)} \sim \int \f{\d^3 n}{n^4}
      \sim \int_{n_{\t{min}}}^\infty \f{\d n}{n^2}
      \sim \f1{n_{\t{min}}},
      \tag{34}
    \end{align*}
    where $n_{\t{min}}$ is defined similarly to $\ell_{\t{min}}$.}

  \step
\item We have replaced the following text (page 12, line 35):

  \red{The values of these counter-terms are fixed by enforcing that
    $G_X$ is equal to the net two-body interaction
    strength. Diagrammatically, this renormalization condition is
    \begin{align*}
      \substack{
        \shrink{
          \begin{tikzpicture}
            \begin{feynman}
              \vertex (v);
              \vertex[above left = of v] (f1);
              \vertex[below left = of v] (f2);
              \vertex[above right = of v] (f3);
              \vertex[below right = of v] (f4);
              \diagram* {
                (f1) -- (v),
                (f2) -- (v),
                (v) -- (f3),
                (v) -- (f4) };
            \end{feynman}
          \end{tikzpicture}}
        \\ \O(G)}
      + \substack{
        \shrink{
          \begin{tikzpicture}
            \begin{feynman}
              \vertex[crossed dot] (v) {};
              \vertex[above left = of v] (f1);
              \vertex[below left = of v] (f2);
              \vertex[above right = of v] (f3);
              \vertex[below right = of v] (f4);
              \diagram* {
                (f1) -- (v),
                (f2) -- (v),
                (v) -- (f3),
                (v) -- (f4) };
            \end{feynman}
          \end{tikzpicture}}
        \\ \O(\tilde G)}
      - \substack{
        \shrink{
          \begin{tikzpicture}
            \begin{feynman}
              \vertex (v1);
              \vertex[above left = of v1] (f1);
              \vertex[below left = of v1] (f2);
              \vertex[right = of v1] (v2);
              \vertex[above right = of v2] (f3);
              \vertex[below right = of v2] (f4);
              \diagram* {
                (f1) -- (v1),
                (f2) -- (v1),
                (v2) -- (f3),
                (v2) -- (f4),
                (v1) --[scalar, half left] (v2),
                (v1) --[scalar, half right] (v2) };
            \end{feynman}
          \end{tikzpicture}}
        \\ \O(G^2)}
      + \cdots = \substack{
        \shrink{
          \begin{tikzpicture}
            \begin{feynman}
              \vertex (v);
              \vertex[above left = of v] (f1);
              \vertex[below left = of v] (f2);
              \vertex[above right = of v] (f3);
              \vertex[below right = of v] (f4);
              \diagram* {
                (f1) -- (v),
                (f2) -- (v),
                (v) -- (f3),
                (v) -- (f4) };
            \end{feynman}
          \end{tikzpicture}}
        \\ \O(G)}.
      \tag{16}
    \end{align*}
    where we represent counter-terms by a crossed dot (i.e.~$\otimes$)
    at a vertex. To leading order in the coupling constants, it
    follows that the counter-terms are second order in the coupling
    constants, i.e.~$\tilde G\sim G^2$. This finding implies that
    counter-terms only come into play at next- to-leading order (NLO)
    in the calculation of any effective $M$-body interaction
    Hamiltonian.}

  by:

  \green{The introduction of counter-terms is merely a formal
    decomposition of the ``bare'' coupling constants $G_X^{\t{bare}}$
    that are used in (15) as $G_X^{\t{bare}} = G_X + \tilde G_X$.  In
    performing such a decomposition, we are free to choose the values
    of $G_X$, which in turn fixes the values of
    $\tilde G_X \equiv G_X^{\t{bare}} - G_X$.  For convenience, we can
    choose the values of $G_X$ to be those of the net effective
    coupling constants on the right-hand side of (15).  Representing
    the new coupling constants $G_X$ by regular vertices and the
    counter-terms $\tilde G_X$ by a crossed dot (i.e.~$\otimes$), this
    choice leads to the {\it renormalization condition}
    \begin{align*}
      \substack{
        \shrink{
          \begin{tikzpicture}
            \begin{feynman}
              \vertex (v);
              \vertex[above left = of v] (f1);
              \vertex[below left = of v] (f2);
              \vertex[above right = of v] (f3);
              \vertex[below right = of v] (f4);
              \diagram* {
                (f1) -- (v),
                (f2) -- (v),
                (v) -- (f3),
                (v) -- (f4) };
            \end{feynman}
          \end{tikzpicture}}
        \\ \O(G)}
      + \substack{
        \shrink{
          \begin{tikzpicture}
            \begin{feynman}
              \vertex[crossed dot] (v) {};
              \vertex[above left = of v] (f1);
              \vertex[below left = of v] (f2);
              \vertex[above right = of v] (f3);
              \vertex[below right = of v] (f4);
              \diagram* {
                (f1) -- (v),
                (f2) -- (v),
                (v) -- (f3),
                (v) -- (f4) };
            \end{feynman}
          \end{tikzpicture}}
        \\ \O(\tilde G)}
      - \substack{
        \shrink{
          \begin{tikzpicture}
            \begin{feynman}
              \vertex (v1);
              \vertex[above left = of v1] (f1);
              \vertex[below left = of v1] (f2);
              \vertex[right = of v1] (v2);
              \vertex[above right = of v2] (f3);
              \vertex[below right = of v2] (f4);
              \diagram* {
                (f1) -- (v1),
                (f2) -- (v1),
                (v2) -- (f3),
                (v2) -- (f4),
                (v1) --[scalar, half left] (v2),
                (v1) --[scalar, half right] (v2) };
            \end{feynman}
          \end{tikzpicture}}
        \\ \O(G^2)}
      + \cdots = \substack{
        \shrink{
          \begin{tikzpicture}
            \begin{feynman}
              \vertex (v);
              \vertex[above left = of v] (f1);
              \vertex[below left = of v] (f2);
              \vertex[above right = of v] (f3);
              \vertex[below right = of v] (f4);
              \diagram* {
                (f1) -- (v),
                (f2) -- (v),
                (v) -- (f3),
                (v) -- (f4) };
            \end{feynman}
          \end{tikzpicture}}
        \\ \O(G)}.
      \tag{16}
    \end{align*}
    This renormalization condition has the benefit of allowing us to
    express effective two-body interactions simply in terms of net
    effective two-body coupling constants, rather than in terms of
    long sums at all order of the bare coupling constants.}

\item The text (page 13, line 17):

  \red{Instead, we must first compute the effective coupling constants
    $G_X^{\t{lattice}}\p{\U}$ in a lattice, which now depend on the
    lattice depth $\U$, and in turn use the effective coupling
    constants to compute interaction energies.}

  now reads:

  \green{The renormalization condition in (16) explicitly fixes $G_X$
    to the net effective coupling constants in any given setting.
    Instead of using free-space coupling constants to compute
    interaction energies in a lattice, we must therefore first compute
    the effective coupling constants $G_X^{\t{lattice}}\p{\U}$, which
    now depend on the lattice depth $\U$, and in turn use these
    effective coupling constants to compute interaction energies.}

\item We have inserted the following paragraph between lines 23 and 24
  on page 13:

  \green{Second, the renormalization condition in (16) implies that
    the counter terms $\tilde G_X$ are second order in the coupling
    constants $G_X$, i.e.~$\tilde G_X\sim G_X^2$.  Although the
    effective Hamiltonian expansions in (13) and (14) are organized in
    powers of the interaction Hamiltonian $H_{\t{int}}$, the couplings
    $G_X$ are the ``small'' parameters in which we can formally
    organize a perturbation theory; more specifically, the formally
    small quantities organizing our perturbation theory are two-body
    ground-state interaction energies (proportional to the couplings
    $G_X$) divided by the spectral gap of the single-atom Hamiltonian
    $H_0$.  If $M$ atoms can only couple through terms represented by
    a $p$-vertex diagrams for $p\ge p_M^{\t{min}}$, then the leading
    order contribution to $M$-body interactions is order
    $p_M^{\t{min}}$ in the couplings $G_X$.  If the same
    $p_M^{\t{min}}$-vertex diagrams involve any counter-terms,
    however, then these diagrams are at least order $p_M^{\t{min}}+1$
    in the couplings $G_X$.  Counter-terms therefore only appear at
    next-to-leading order (NLO) in the calculation of effective
    $M$-body interactions.}

\item The following text (page 13, line 7):

  \red{Before moving on to consider effective three-body interactions,
    there are two comments we must make concerning the result in
    (17).}

  now reads:

  \green{Before moving on to consider effective three-body
    interactions, there are a few comments we must make concerning
    renormalization and the result in (17).}

\end{enumerate}

\section*{Revisions in response to Referee 2}

\setcounter{point}{1}
\begin{enumerate}[label=(R2.\arabic{point}.\arabic*)]
\item
\end{enumerate}

\section*{Miscellaneous revisions}

\begin{enumerate}[label=(R3.\arabic*)]
\item Page 3, line 24 of our original manuscript previously said:

  \red{Last year (2017), a new generation of OLCs became operational
    at JILA}

  Reflecting the fact that it is now 2019, the new version of the
  manuscript says:

  \green{More recently (2017), a new generation of OLCs became
    operational at JILA}

\item Page 5, line 40 of our original manuscript previously said:

  \red{motional state $n\in\mathbb{N}_0^3$ with nuclear spin
    $\mu\in\set{-I,I+1,\cdots,I}$}

  This text contains a sign error; in the version of our manuscript,
  this text reads:

  \green{motional state $n\in\mathbb{N}_0^3$ with nuclear spin
    $\mu\in\set{-I,-I+1,\cdots,I}$}

\item Page 5, lines 43-44 of our original manuscript previously said:

  \red{In a harmonic trap approximation we would have
    $E_n=\p{3/2+n_x+n_y+n_z}\omega$}

  For clarity, we have modified this text so that all subscripts
  denoting free variables (e.g.~the $n$ in $E_n$, as $n$ can here be
  any element of $\mathbb{N}_0^3$) are italic, while all subscripts
  denoting fixed symbols (e.g.~the $x$ in $n_x$, as here $x$ denotes a
  fixed spatial axis) are in roman font.  The new text thus reads:

  \green{In a harmonic trap approximation we would have
    $E_n=\p{3/2+n_\x+n_\y+n_\z}\omega$}

  We have similarly replaced all appearances of $m_A$ (denoting the
  mass of a single atom) with $m_{\text{A}}$; all $E_R$ (denoting the
  lattice photon recoil energy) with $E_{\t{R}}$; all
  $k_L,k_L^x,k_L^y,k_L^z$ (denoting the wavenumber of the lattice
  laser and its projection onto three lattice axes) by
  $k_{\text{L}}, k_{\text{L}}^\x, k_{\text{L}}^\y, k_{\text{L}}^\z$;
  all $\U_x,\U_y,\U_z$ (denoting lattice depths along lattice axes)
  respectively by $\U_\x,\U_\y,\U_\z$; and all
  $T_\mu^z,T_\mu^x,S_{\N X}^z,S_{\N X}^x$ (denoting single-atom and
  many-body pseudospin operators) respectively by
  $T_\mu^\z,T_\mu^\x,S_{\N X}^\z,S_{\N X}^\x$.

  Finally, we have fixed a few discrepancies in our general convention
  that multiple subscripts on a given mathematical symbol should only
  be separated by a comma if at least one of these subscripts has a
  fixed value (as e.g.~in $K^{k\ell}_{mn}$, $K^{\ell,3}_{2,5}$,
  $\c_{\mu n}$, and $\c_{\mu,\g}$).  To this end, we have removed the
  commas on the subscripts ``$M,X$''; ``$N,\pm$''; and ``$\N,\pm$''
  whenever they appear.

\item On page 7, line 17 we previously had the text:

  \red{The existence of such a gauge is guaranteed by the analytic
    properties of Wannier orbitals $\phi_{in}$ [42].}

  With the exception of this sentence, our manuscript has generally
  reserved the word ``orbital'' to refer to electronic degrees of
  freedom of atoms, and ``motional state'' to refer to atoms' spatial
  degrees of freedom.  To avoid potential confusion, we have therefore
  modified the above sentence to read:

  \green{The existence of such a gauge is guaranteed by the analytic
    properties of Wannier wavefunctions $\phi_{in}$ [42].}

\item The last two lines of page 10 and line 26 of page 11 previously
  said:

  \red{Effective $M$-body interaction terms at order $p$ in
    $H_{\t{int}}$ can be represented by directed, weakly connected
    graphs (i.e.~which are connected when ignoring edge direction)
    containing $p$ internal vertices.}

  Though the final expressions in our perturbation theory can indeed
  be represented by weakly connected graphs, intermediate terms that
  cancel out may be represented by graphs that are not connected, as
  e.g.~in Eq.~(35) on page 17.  Furthermore, while the meaning of an
  ``internal vertex'' is likely clear from context, it is better for
  us to define it explicitly.  For clarity, we have therefore modified
  the above text to read:

  \green{Effective $M$-body interaction terms at order $p$ in
    $H_{\t{int}}$ can be represented by directed graphs containing $p$
    vertices with degree greater than one, which we call {\it internal
      vervices}.}

\item To clean up repetitive wording, we have replaced the sentence
  (page 12, line 45):

  \red{To leading order in the coupling constants, it follows that the
    counter-terms are second order in the coupling constants,
    i.e. $\tilde G\sim G^2$.}

  with:

  \green{To leading order, it follows that the counter-terms are
    second order in the coupling constants, i.e.~$\tilde G\sim G^2$.}

\item We fixed an inconsistency in our convention for the ordering of
  operators associated with the coupling symbol
  $G^{\mu q;\nu r}_{\rho s;\sigma t}$.  The text which previously read
  (page 14, line 47):

  \red{for more general
    $\p{\mu,q}+\p{\nu,r}\leftrightarrow\p{\rho,s}+\p{\sigma,t}$
    coupling induced by terms proportional to
    $\c_{\mu q}^\dag \c_{\nu r}^\dag \c_{\sigma t} \c_{\rho s}$.}

  now reads:

  \green{for more general
    $\p{\mu,q}+\p{\nu,r}\leftrightarrow\p{\rho,s}+\p{\sigma,t}$
    coupling induced by terms proportional to
    $\c_{\rho s}^\dag \c_{\sigma t}^\dag \c_{\nu r} \c_{\mu q}$.}

  Both examples are ``correct'' due to the fact that
  $G^{\mu q;\nu r}_{\rho s;\sigma t}$ is self-adjoint, but the latter
  ordering convention is used throughout the rest of the manuscript:
  $G^{\mu q;\nu r}_{\rho s;\sigma t}$ is gives the coupling constant
  for destroying two atoms in state $\p{\mu q}+\p{\nu r}$, and
  creating them in $\p{\rho s}+\p{\sigma t}$.

\item In Eq.~(21), we have fixed an error in the nuclear spin indices
  ($\mu,\nu$) on the two rightmost operators of the middle term, such
  that (page 14, line 53):

  \red{\begin{align*}
      G^{\mu q;\nu r}_{\nu s;\mu t}
      \c_{\nu s}^\dag \c_{\mu t}^\dag \c_{\nu r} \c_{\mu q}
      = -G^{qr}_{ts} \c_{\nu s}^\dag \c_{\mu t}^\dag \c_{\mu r} \c_{\nu q}
      = G^{qr}_{ts} \c_{\mu t}^\dag \c_{\nu s}^\dag \c_{\nu r} \c_{\mu q}.
      \tag{21}
    \end{align*}}

  now reads:

  \green{\begin{align*}
      G^{\mu q;\nu r}_{\nu s;\mu t}
      \c_{\nu s}^\dag \c_{\mu t}^\dag \c_{\nu r} \c_{\mu q}
      = -G^{qr}_{ts} \c_{\nu s}^\dag \c_{\mu t}^\dag \c_{\nu r} \c_{\mu q}
      = G^{qr}_{ts} \c_{\mu t}^\dag \c_{\nu s}^\dag \c_{\nu r} \c_{\mu q}.
      \tag{21}
    \end{align*}}

\item To make our first appendix more self-contained, we have defined
  the projector $\P_0$ used therein, such that the text (page 30, line
  38):

  \red{Letting $\Q_0\equiv1-\P_0$ denote the projector onto $\E_0$ and
    $X$ ...}

  now reads:

  \green{Letting $\P_0$ denote the projector onto $\G_0$,
    $\Q_0\equiv\1-\P_0$ denote the projector onto $\E_0$, and $X$ ...}

\item For clarity, we have replaced a few instances of the text
  ``\red{single-particle}'' by ``\green{non-interacting}'' when
  referring to motional ground states of non-interacting particles;
  specifically, we have done so on page 9, line 50; on page 13, line
  5; and on page 35, line 26.

\item Reference 34 of the original manuscript has been published since
  the original submission of our manuscript.  We have updated this
  reference accordingly in the bibliography of our revised manuscript.

\end{enumerate}

\end{document}

### 2019-02-13 multi_body_interactions/papers/theory/reviews_NJP/referee_response_1.tex
changed instance of 'single-atom' to 'non-interacting', and added one revision

@@
     lattice site [34].  At ultracold temperatures well below the
-    single-atom motional excitation energies $\Delta_n\equiv E_n-E_0$
-    for $n>0$, atoms only occupy their motional ground state in the
-    lattice.  For this reason, it is common to map the description of
-    these atoms onto a single-band Hubbard model that captures all
-    dynamics within the subspace of motional ground states of
-    non-interacting atoms, i.e.~with single-atom wavefunctions
-    $\phi_{i,0}$.  Interactions, however, modify atoms' motional
-    ground-state wavefuctions.  The true motional ground state of a
-    collection of interacting atoms is then an admixture of the
-    non-interacting motional eigenstates, and a naive Hubbard model
-    that assumes atomic wavefunctions $\phi_{i,0}$ will fail to
+    non-interacting motional excitation energies
+    $\Delta_n\equiv E_n-E_0$ for $n>0$, atoms only occupy their
+    motional ground state in the lattice.  For this reason, it is
+    common to map the description of these atoms onto a single-band
+    Hubbard model that captures all dynamics within the subspace of
+    motional ground states of non-interacting atoms, i.e.~with
+    wavefunctions $\phi_{i,0}$.  Interactions, however, modify atoms'
+    motional ground-state wavefuctions.  The true motional ground
+    state of a collection of interacting atoms is then an admixture of
+    the non-interacting motional eigenstates, and a naive Hubbard
+    model that assumes atomic wavefunctions $\phi_{i,0}$ will fail to
     reproduce the interacting atoms' orbital excitation spectrum.

### 2019-02-13 multi_body_interactions/papers/theory/paper.tex
changed instance of 'single-atom' to 'non-interacting', and added one revision

@@
 site \cite{goban2018emergence}.  At ultracold temperatures well below
-the single-atom motional excitation energies $\Delta_n\equiv E_n-E_0$
-for $n>0$, atoms only occupy their motional ground state in the
-lattice.  For this reason, it is common to map the description of
-these atoms onto a single-band Hubbard model that captures all
-dynamics within the subspace of motional ground states of
-non-interacting atoms, i.e.~with single-atom wavefunctions
-$\phi_{i,0}$.  Interactions, however, modify atoms' motional
-ground-state wavefuctions.  The true motional ground state of a
-collection of interacting atoms is then an admixture of the
-non-interacting motional eigenstates, and a naive Hubbard model that
-assumes atomic wavefunctions $\phi_{i,0}$ will fail to reproduce the
-interacting atoms' orbital excitation spectrum.  Formally, corrections
-to the spectrum of interacting atoms can be accounted for by a
-perturbative treatment of far-off-resonant terms in the interaction
-Hamiltonian of \eqref{eq:H_int} that create atoms in excited motional
-states,
+the non-interacting motional excitation energies
+$\Delta_n\equiv E_n-E_0$ for $n>0$, atoms only occupy their motional
+ground state in the lattice.  For this reason, it is common to map the
+description of these atoms onto a single-band Hubbard model that
+captures all dynamics within the subspace of motional ground states of
+non-interacting atoms, i.e.~with wavefunctions $\phi_{i,0}$.
+Interactions, however, modify atoms' motional ground-state
+wavefuctions.  The true motional ground state of a collection of
+interacting atoms is then an admixture of the non-interacting motional
+eigenstates, and a naive Hubbard model that assumes atomic
+wavefunctions $\phi_{i,0}$ will fail to reproduce the interacting
+atoms' orbital excitation spectrum.  Formally, corrections to the
+spectrum of interacting atoms can be accounted for by a perturbative
+treatment of far-off-resonant terms in the interaction Hamiltonian of
+\eqref{eq:H_int} that create atoms in excited motional states,
 e.g.~$\sim\c_{n\mu s}^\dag \c_{0,\nu t}^\dag \c_{0,\nu r} \c_{0,\mu
@@
     as a visual guide.  Resonance peaks are identified by the
-    many-body orbital states which are excited at the peak.
-    \label{fig:sweep}}
+    many-body orbital states which are excited at the
+    peak.}
+  \label{fig:sweep}
 \end{figure}
@@
 We now consider samples of ${}^{87}$Sr atoms in a uniform mixture of
-nuclear spins $\mu\in\set{-I,\cdots,I}$ prepared in motional ground
-states of a rectangular lattice with depths
+nuclear spins $\mu\in\set{-I,-I+1,\cdots,I}$ prepared in motional
+ground states of a rectangular lattice with depths
 $\U=\p{\U_\x,\U_\y,\U_\z}=\p{41,55,69} E_{\t{R}}$, where

### 2019-02-13 multi_body_interactions/papers/theory/reviews_NJP/revision_summary.tex
changed instance of 'single-atom' to 'non-interacting', and added one revision

@@
 
+\item In figure 2 (page 22), all multiplicities (i.e.~values on the
+  vertical axes) are integer-valued.  We have therefore adjusted the
+  vertical axis on the top left panel of this figure, such that now
+  only integer values are marked on this axis.
+
+\item For slight clarification, we have changed the text (page 22,
+  line 51):
+
+  \red{... nuclear spins $\mu\in\set{-I,\cdots,I}$ ...}
+
+  to:
+
+  \green{... nuclear spins $\mu\in\set{-I,-I+1,\cdots,I}$ ...}
+
 \item To make our first appendix more self-contained, we have defined
@@
 \item For clarity, we have replaced a few instances of the text
-  ``\red{single-particle}'' by ``\green{non-interacting}'' when
-  referring to motional ground states of non-interacting particles;
-  specifically, we have done so on page 9, line 50; on page 13, line
-  5; and on page 35, line 26.
+  ``\red{single-particle}'' or ``\red{single-atom}'' by
+  ``\green{non-interacting}'' when referring to motional ground states
+  of non-interacting particles; specifically, we have done so on page
+  9, line 50; on page 11, line 22; on page 13, line 5; and on page 35,
+  line 26.
 

### 2019-02-13 multi_body_interactions/papers/theory/reviews_NJP/referee_response_2.tex
added some responses to comments from referee 2 from NJP

\documentclass[preprint]{revtex4-1}

% linking references
\usepackage{hyperref}
\hypersetup{
  breaklinks=true,
  colorlinks=true,
  linkcolor=blue,
  urlcolor=cyan,
}

% general physics / math packages and commands
\usepackage{physics,amsmath,amssymb,braket,dsfont}
\renewcommand{\t}{\text} % text in math mode
\newcommand{\f}{\dfrac} % shorthand for fractions
\newcommand{\p}[1]{\left(#1\right)} % parenthesis
\renewcommand{\sp}[1]{\left[#1\right]} % square parenthesis
\renewcommand{\set}[1]{\left\{#1\right\}} % curly parenthesis
\newcommand{\bk}{\braket} % shorthand for braket

\renewcommand{\d}{\text{d}}
\newcommand{\g}{\text{g}}
\newcommand{\e}{\text{e}}
\newcommand{\x}{\text{x}}
\newcommand{\y}{\text{y}}
\newcommand{\z}{\text{z}}

\renewcommand{\c}{\hat{c}}
\newcommand{\n}{\hat{n}}

\newcommand{\A}{\mathcal{A}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\G}{\mathcal{G}}
\renewcommand{\H}{\mathcal{H}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\K}{\mathcal{K}}
\renewcommand{\L}{\mathcal{L}}
\newcommand{\M}{\mathcal{M}}
\newcommand{\N}{\mathcal{N}}
\renewcommand{\O}{\mathcal{O}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\Q}{\mathcal{Q}}
\renewcommand{\S}{\mathcal{S}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\1}{\mathds{1}}

% inline lists
\usepackage[inline]{enumitem}
\setlist[enumerate]{leftmargin=*} % nice margins for enumerate

% for feynman diagrams
\usepackage{tikz,tikz-feynman}
\tikzset{
  baseline = (current bounding box.center)
}
\tikzfeynmanset{
  compat = 1.1.0,
  every feynman = {/tikzfeynman/small}
}
\newcommand{\shrink}[1]{\scalebox{0.8}{#1}} % for smaller diagrams

% color definitions (used in a figure)
\usepackage{xcolor}
\newcommand{\blue}[1]{\textcolor{blue}{#1}}
\newcommand{\red}[1]{\textcolor{red}{#1}}
\newcommand{\green}[1]{\textcolor{green}{#1}}
\definecolor{lightblue}{RGB}{31,119,180}
\definecolor{orange}{RGB}{255,127,14}
\definecolor{green}{RGB}{44,160,44}
\definecolor{lightred}{RGB}{214,39,40}

% for strikeout text
% normalem included to prevent underlining titles in the bibliography
\usepackage[normalem]{ulem}

\begin{document}

\section*{Referee 2}

We thank the referee for taking the time to carefully read our
manuscript, and for providing thoughtful and constructive feedback.
Below, we address the comments and concerns brought up by the referee.
We hope that the referee finds our responses and revisions
satisfactory, deeming our manuscript acceptable for publication in New
Journal of Physics.

Excerpts from the referee reports are written in \blue{blue}, excerpts
from our original manuscript are written in \red{red}, and excerpts
from the revised version of our manuscript is written in
\green{green}.  All page, line, equation, and reference numbers
generally refer to those in our original manuscript, unless explicitly
stated otherwise.

\begin{enumerate}
\item Concerning:

  \blue{Largely motivated by the experimental spectroscopy data on
    87Sr atoms, the authors of manuscript NJP-109533 have studied
    effective few-body interactions in mesoscopic Sr atomic clusters
    confined by optical lattice potentials and applied the effective
    Hamiltonian to characterize the spectra of interacting Sr atoms.
    The authors have also compared the result of theory with data and
    shown general consistency.}

  \blue{The main results are summarized in section III and IV.  In
    section II, the authors presented the detailed discussions on the
    effective interactions via the standard Schrieffer-Wolff
    transformation. The low energy manifold where the orbital motion
    is fixed in the ground state is separated from excited manifolds
    involving excited orbitals. The low energy physics therefore is
    purely driven by the internal SU(N) spin dynamics and the excited
    orbitals will further induce few-body interaction between SU(N)
    spins.  These can be systematically studied at least in the weakly
    interacting systems or in the perturbative limit where the single
    particle orbital motion is well gapped and Schrieffer-Wolff
    transformation is applicable. In section IV, the authors benched
    their numerical results against the data on 87Sr and emphasized
    the situation where the induced effective interactions might play
    a substantial role.}

  \blue{I find the article is well written and results have been
    carefully organized and are well presented.  Although at
    qualitatively level, these higher order effects are relatively
    small (not surprising as perturbative effects) and do not modify
    the results in a drastic way, they do offer better agreements with
    experimental data. And in the extreme cases in Fig.5 (3+, 4+, 5+
    lines), numerical values of these effects o(G\^{}3) can be as
    larger as 10-20\% (to my bare eyes) although in this limit it
    becomes less clear whether the contributions not included are also
    substantial.}

  \blue{A few suggestions the authors can take into account to improve
    the presentation.}

  We thank the referee for their summary of our work, and for
  remarking that our ``article is well written'', and that our
  ``results have been carefully organized are well presented''.  We
  hope that the revisions we outline below address the referee's
  suggestions to a satisfactory degree.

\item Concerning:

  \blue{1) In the caption of Fig 1, the authors shall include the
    dashed solid line definitions to make it self-contained.}

  We thank the referee for the suggestion to make figure 1 more
  self-contained.  Following this suggestion, we have added the
  definition of solid and dashed lines to the caption of figure 1, as
  well as a sentence describing how to translate a diagram into a
  sequence of operators.  In total, we have replaced the text (page
  11, line 13):

  \red{... $n>0$ and $\c_{\mu s}\equiv\c_{0,\mu s}$.  For the sake
    presentation, ...}

  by:

  \green{... $n>0$ and $\c_{\mu s}\equiv\c_{0,\mu s}$.  Diagrams are
    read from left to right to construct a sequence of operators from
    right to left.  Solid (dashed) lines represent field operators
    acting on the lowest (arbitrary) motional states.  For the sake
    presentation, ...}

\item Concerning the closely related comments/questions::

  \blue{2) Section IV, the author shall add discussions about why some
    states are more affected than the others by the 4-body interaction
    etc.  Data are clearly presented and well compared with
    experiments but there seems to be a lack of central message. What
    is the main lesson? After all, they are high order effects so
    naively and logically one would say they play little role in the
    system. But on other hand, numerically in some cases they appear
    to be amplified up to 20\%, not so small anymore.  Is it
    accidental for 3+, 4+, 5+ states or there are some more
    fundamental reasons for that?}

  \blue{3) Related to 2), when the correction becomes 10-20\%, shall
    we expect that higher order effects of order of o(G\^{}4) etc also
    become important?}

  We thank the referee for their careful reading of our results in
  section IV, and for pointing out that our manuscript would greatly
  benefit from a deeper discussion of figures 4 and 5.  The referee
  asks questions that follow naturally from an inspection of these
  figures, which our manuscript should address directly.

  Indeed, there is a good reason why multi-body interaction energy
  corrections are more prominent for the symmetric ($3+$, $4+$, $5+$)
  states than for the asymmetric ($3-$, $4-$, $5-$) ones.
  Specifically, there is a competition between contributions of
  opposite sign to the interaction energy of asymmetric states.
  Concerning higher-order effects, it is not obvious when exactly
  further corrections ($\sim G^4$) should become relevant.  Computing
  higher-order corrections requires more systematic methods than those
  used in our manuscript, and we are simply unsure whether considering
  these corrections is worthwhile without first performing a more
  detailed analysis of systematic errors in both theory and
  experiment.

  To address these questions and add a more pointed discussion of the
  lessons from figures 4 and 5, we have rearranged, rewritten, and
  added new text to the following two paragraphs (page 23, line 54):

  \red{Identifying peaks in excitation spectra such as in figure 3
    constitutes a measurement of many-body excitation energies, which
    was performed in ref.~[34] to detect signatures of effective
    multi-body interactions.  Figure 4 shows a comparison between
    experimental measurements of the many-body excitation energies
    $\Delta_{NX}$ for $N\in\set{3,4,5}$ at various lattice depths
    $\U$, and the corresponding values of $\Delta_{NX}$ predicted by
    the low-energy effective theory at different orders in the
    coupling constants.  A summary of the data in figure 4 is provided
    in figure 5.  These figures show a clear improvement of the
    low-energy effective theory we have developed with increasing
    order in the coupling constants, in particular demonstrating the
    need to consider multi-body interactions in a single-band model of
    the atoms without motional excitations.}

  \red{We expect that the dominant source of error in the third-order
    effective theory comes from the fact that we neglect the
    inter-site matrix elements of all Hamiltonians.  This source of
    error is discussed in Appendix E, and leads to theoretical
    uncertainties represented by error bars in figure 4.  As these
    uncertainties are sometimes unable to account for the residual
    disagreement between experimental measurements and the effective
    theory, better agreement requires either a better understanding of
    experimental procedures and error, a better understanding of
    theoretical uncertainties, or consideration of yet higher-order
    terms in the effective theory.  Due to a combinatorial explosion
    in the number of diagrams which appear at increasing perturbative
    orders of the effective theory, we need more systematic methods to
    compute effective multi-body Hamiltonians at fourth order; we
    therefore leave this calculation for future work.}

  which now read:

  \green{Identifying peaks in excitation spectra such as in figure 3
    constitutes a measurement of many-body excitation energies, which
    was performed in ref.~[34] to detect signatures of effective
    multi-body interactions.  Figure 4 shows a comparison between (i)
    experimental measurements of the many-body excitation energies
    $\Delta_{NX}$ for all $\p{N,X}\in\set{3,4,5}\times\set{+,-}$ at
    various lattice depths $\U$, and (ii) the corresponding values of
    $\Delta_{NX}$ predicted by the low-energy effective theory at
    different orders in the coupling constants.  A known source of
    error in our effective theory comes from neglecting the inter-site
    matrix elements of all Hamiltonians.  This error is discussed in
    Appendix E, and leads to theoretical uncertainties represented by
    error bars on the $\O\p{G^3}$ theory in figure 4.  A summary of
    figure 4 is provided in figure 5.  We note that many-body
    interaction energy shifts are smaller for asymmetric ($-$) states
    than symmetric ($+$) ones due to the competition between
    contributions of opposite sign in the asymmetric case (see rows 2
    and 3 of table I).}

  \green{The results in figures 4 and 5 highlight a few important
    points about ultracold, high-density ${}^{87}$Sr experiments and
    our low-energy effective theory.  First, these experiments exhibit
    clear signatures of multi-body interactions, as evidenced by a
    stark disagreement between the observed many-body excitation
    energies $\Delta_{NX}$ and those that are predicted by the
    two-body $\O\p{G}$ theory.  Multi-body interactions are thus
    crucial for understanding these high-density experiments in the
    context of a single-band Hubbard model, which naturally arises in
    the zero-temperature limit when all atoms occupy their motional
    ground state.  Second, the inter-atomic interactions in these
    experiments are strong enough to require going beyond leading
    order for the description of multi-body interactions in the
    low-energy effective theory.  As experiments begin to work with
    more atoms and shallower lattices, therefore, reliably predicting
    interaction energies may require going to yet higher orders in
    perturbation theory.  Due to a combinatorial explosion of the
    number of diagrams which appear at increasing orders in the
    effective theory, however, we need more systematic methods to
    compute effective multi-body Hamiltonians at fourth order.  In any
    case, we are agnostic as to whether such a calculation would
    provide better agreement between experiment and theory without
    first performing a more detailed analysis of systematic errors.}

  We hope that this revision has clarified the questions brought up by
  the referee, and provided a central message for the results in
  figures 4 and 5.

\item Concerning:

  \blue{4) At a phenomenological level, the author shall describe, in
    connection to the experiment, the range of the perturbation
    parameter [ a\_X K/Delta] that measures the two-particle
    interaction energy. For Sr atoms with typical background
    scattering lengths, this is usually small but does depend on the
    lattice depth quite sensitively.  The authors shall provide the
    typical numerical range of this dimensionless interaction energy
    for the optical lattices (up to 70E\_R in Fig 5) they studied (and
    experimentally studied). This will help readers to better
    appreciate their numerical results (surprising or anticipated).}

\end{enumerate}

\end{document}

### 2019-02-13 multi_body_interactions/papers/theory/paper.tex
added some responses to comments from referee 2 from NJP

@@
     corresponding three-body interaction term in $H_{\t{int}}^{(2)}$,
-    with $n>0$ and $\c_{\mu s}\equiv\c_{0,\mu s}$.  For the sake of
+    with $n>0$ and $\c_{\mu s}\equiv\c_{0,\mu s}$.  Diagrams are read
+    from left to right to construct a sequence of operators from right
+    to left.  Solid (dashed) lines represent field operators acting on
+    the lowest (arbitrary) motional states.  For the sake of
     presentation, this diagram has colors associated with nuclear spin
@@
     understanding that the low-energy effective theory implicitly
-    addresses only single-atom motional ground states.
-    \label{fig:diagram}}
+    addresses only non-interacting motional ground states.}
+  \label{fig:diagram}
 \end{figure}
@@
     single-site occupation numbers $N$ which are achievable in current
-    ${}^{87}$Sr experiments.  \label{fig:coefficients}}
+    ${}^{87}$Sr experiments.}
+  \label{fig:coefficients}
 \end{figure}
@@
     states in the low-energy effective theory (see Appendix
-    \ref{sec:error}).  \label{fig:shifts}}
+    \ref{sec:error}).}
+  \label{fig:shifts}
 \end{figure}
@@
     in figure \ref{fig:shifts}, retrieved from
-    ref.~\cite{goban2018emergence}.  \label{fig:shifts_summary}}
+    ref.~\cite{goban2018emergence}.}
+  \label{fig:shifts_summary}
 \end{figure}
@@
 detect signatures of effective multi-body interactions.  Figure
-\ref{fig:shifts} shows a comparison between experimental measurements
-of the many-body excitation energies $\Delta_{NX}$ for
-$N\in\set{3,4,5}$ at various lattice depths $\U$, and the
-corresponding values of $\Delta_{NX}$ predicted by the low-energy
-effective theory at different orders in the coupling constants.  A
-summary of the data in figure \ref{fig:shifts} is provided in figure
-\ref{fig:shifts_summary}.  These figures show a clear improvement of
-the low-energy effective theory we have developed with increasing
-order in the coupling constants, in particular demonstrating the need
-to consider multi-body interactions in a single-band model of the
-atoms without motional excitations.
-
-We expect that the dominant source of error in the third-order
-effective theory comes from the fact that we neglect the inter-site
-matrix elements of all Hamiltonians.  This source of error is
+\ref{fig:shifts} shows a comparison between
+\begin{enumerate*}
+\item experimental measurements of the many-body excitation energies
+  $\Delta_{NX}$ for all $\p{N,X}\in\set{3,4,5}\times\set{+,-}$ at
+  various lattice depths $\U$, and
+\item the corresponding values of $\Delta_{NX}$ predicted by the
+  low-energy effective theory at different orders in the coupling
+  constants.
+\end{enumerate*}
+A known source of error in our effective theory comes from neglecting
+the inter-site matrix elements of all Hamiltonians.  This error is
 discussed in Appendix \ref{sec:error}, and leads to theoretical
-uncertainties represented by error bars in figure \ref{fig:shifts}.
-As these uncertainties are sometimes unable to account for the
-residual disagreement between experimental measurements and the
-effective theory, better agreement requires either a better
-understanding of experimental procedures and error, a better
-understanding of theoretical uncertainties, or consideration of yet
-higher-order terms in the effective theory.  Due to a combinatorial
-explosion in the number of diagrams which appear at increasing
-perturbative orders of the effective theory, we need more systematic
-methods to compute effective multi-body Hamiltonians at fourth order;
-we therefore leave this calculation for future work.
+uncertainties represented by error bars on the $\O\p{G^3}$ theory in
+figure \ref{fig:shifts}.  A summary of figure \ref{fig:shifts} is
+provided in figure \ref{fig:shifts_summary}.  We note that many-body
+interaction energy shifts are smaller for asymmetric ($-$) states than
+symmetric ($+$) ones due to the competition between contributions of
+opposite sign in the asymmetric case (see rows 2 and 3 of table
+\ref{tab:eigen}).
+
+The results in figures \ref{fig:shifts} and \ref{fig:shifts_summary}
+highlight a few important points about ultracold, high-density
+${}^{87}$Sr experiments and our low-energy effective theory.  First,
+these experiments exhibit clear signatures of multi-body interactions,
+as evidenced by a stark disagreement between the observed many-body
+excitation energies $\Delta_{NX}$ and those that are predicted by the
+two-body $\O\p{G}$ theory.  Multi-body interactions are thus crucial
+for understanding these high-density experiments in the context of a
+single-band Hubbard model, which naturally arises in the
+zero-temperature limit when all atoms occupy their motional ground
+state.  Second, the inter-atomic interactions in these experiments are
+strong enough to require going beyond leading order for the
+description of multi-body interactions in the low-energy effective
+theory.  As experiments begin to work with more atoms and shallower
+lattices, therefore, reliably predicting interaction energies may
+require going to yet higher orders in perturbation theory.  Due to a
+combinatorial explosion of the number of diagrams which appear at
+increasing orders in the effective theory, however, we need more
+systematic methods to compute effective multi-body Hamiltonians at
+fourth order.  In any case, we are agnostic as to whether such a
+calculation would provide better agreement between experiment and
+theory without first performing a detailed analysis of systematic
+errors.
 
@@
     $\P_{NX}$, defined in \eqref{eq:P_NX}, is a projector onto the
-    space of the $N$-body orbital states $\ket{\N X}$.
-    \label{fig:time}}
+    space of the $N$-body orbital states $\ket{\N X}$.}
+  \label{fig:time}
 \end{figure}

### 2019-02-14 multi_body_interactions/papers/theory/reviews_NJP/referee_response_1.tex
added last of the revisions to the paper (pre-markup by AMR)

@@
-\documentclass[preprint]{revtex4-1}
+\documentclass[preprint,showkeys,nofootinbib]{revtex4-1}
 
@@
 \newcommand{\U}{\mathcal{U}}
+
 \newcommand{\1}{\mathds{1}}
+
+\newcommand{\mA}{m_{\text{A}}} % symbol for the mass of an atom
+
+% "left vector" arrow; requires tikz package
+\newcommand{\lvec}[1]
+{\reflectbox{\ensuremath{\vec{\reflectbox{\ensuremath{#1}}}}}}
+
+% figures
+\usepackage{graphicx} % for figures
+\usepackage{grffile} % help latex properly identify figure extensions
+\graphicspath{{./figures/}} % set path for all figures
+\usepackage[caption=false]{subfig} % subfigures (via \subfloat[]{})
 
@@
 \usepackage[inline]{enumitem}
-\setlist[enumerate]{leftmargin=*} % nice margins for enumerate
 
@@
 \usepackage{xcolor}
-\newcommand{\blue}[1]{\textcolor{blue}{#1}}
-\newcommand{\red}[1]{\textcolor{red}{#1}}
-\newcommand{\green}[1]{\textcolor{green}{#1}}
 \definecolor{lightblue}{RGB}{31,119,180}
@@
 
+% proper coloring inside math environment
+\makeatletter
+\def\mathcolor#1#{\@mathcolor{#1}}
+\def\@mathcolor#1#2#3{
+  \protect\leavevmode
+  \begingroup
+    \color#1{#2}#3
+  \endgroup
+}
+\makeatother
+\newcommand{\bmu}{\mathcolor{lightblue}{\mu}}
+\newcommand{\onu}{\mathcolor{orange}{\nu}}
+\newcommand{\grho}{\mathcolor{green}{\rho}}
+\newcommand{\re}{\mathcolor{lightred}{\text{e}}}
+
+% leave a note in the text, visible in the compiled document
+\newcommand{\note}[1]{\textcolor{red}{#1}}
+
 % for strikeout text
@@
 \usepackage[normalem]{ulem}
+
+\newcommand{\blue}[1]{\textcolor{blue}{#1}}
+\newcommand{\red}[1]{\textcolor{red}{#1}}
+\newcommand{\green}[1]{\textcolor{green}{#1}}
 
@@
   new equations in response to point \ref{pt:divergence} below.  We
-  have made the same change to equation (F1), which is included in
-  Appendix F for quick reference.
+  have made the same change to Eq.~(F1), which is included in Appendix
+  F for quick reference.  Appendix F and Eq.~(F1) are respectively
+  Appendix G and Eq.~(G1) in the revised version of our manuscript, as
+  we have added a new appendix in one of our responses to referee 2.
 
@@
     $G_X$) divided by the spectral gap of the single-atom Hamiltonian
-    $H_0$.  If $M$ atoms can only couple through terms represented by
-    a $p$-vertex diagrams for $p\ge p_M^{\t{min}}$, then the leading
-    order contribution to $M$-body interactions is order
-    $p_M^{\t{min}}$ in the couplings $G_X$.  If the same
-    $p_M^{\t{min}}$-vertex diagrams involve any counter-terms,
-    however, then these diagrams are at least order $p_M^{\t{min}}+1$
-    in the couplings $G_X$.  Counter-terms therefore only appear at
-    next-to-leading order (NLO) in the calculation of effective
-    $M$-body interactions.}
-
-  To reflect this added paragraph, we have modified the text on page
-  13, line 7:
+    $H_0$ (see Appendix F).  If $M$ atoms can only couple through
+    terms represented by a $p$-vertex diagrams for
+    $p\ge p_M^{\t{min}}$, then the leading order contribution to
+    $M$-body interactions is order $p_M^{\t{min}}$ in the couplings
+    $G_X$.  If the same $p_M^{\t{min}}$-vertex diagrams involve any
+    counter-terms, however, then these diagrams are at least order
+    $p_M^{\t{min}}+1$ in the couplings $G_X$.  Counter-terms therefore
+    only appear at next-to-leading order (NLO) in the calculation of
+    effective $M$-body interactions.}
+
+  Note that Appendix F referenced above is a new appendix we have
+  added in our response to referee 2.  To reflect the addition of the
+  above paragraph, we have modified the text on page 13, line 7:
 

### 2019-02-14 multi_body_interactions/papers/theory/reviews_NJP/referee_response_2.tex
added last of the revisions to the paper (pre-markup by AMR)

@@
 
+  \label{pt:figs}
+
   We thank the referee for their careful reading of our results in
@@
   asks questions that follow naturally from an inspection of these
-  figures, which our manuscript should address directly.
+  figures; our manuscript should address these questions directly.
 
@@
   lessons from figures 4 and 5, we have rearranged, rewritten, and
-  added new text to the following two paragraphs (page 23, line 54):
+  added new text to the following two paragraphs about these figures
+  (page 23, line 54):
 
@@
     $\Delta_{NX}$ for all $\p{N,X}\in\set{3,4,5}\times\set{+,-}$ at
-    various lattice depths $\U$, and (ii) the corresponding values of
-    $\Delta_{NX}$ predicted by the low-energy effective theory at
-    different orders in the coupling constants.  A known source of
-    error in our effective theory comes from neglecting the inter-site
-    matrix elements of all Hamiltonians.  This error is discussed in
-    Appendix E, and leads to theoretical uncertainties represented by
-    error bars on the $\O\p{G^3}$ theory in figure 4.  A summary of
-    figure 4 is provided in figure 5.  We note that many-body
-    interaction energy shifts are smaller for asymmetric ($-$) states
-    than symmetric ($+$) ones due to the competition between
-    contributions of opposite sign in the asymmetric case (see rows 2
-    and 3 of table I).}
+    various mean lattice depths $\U$, and (ii) the corresponding
+    values of $\Delta_{NX}$ predicted by the low-energy effective
+    theory at different orders in the coupling constants.  A known
+    source of error in our effective theory comes from neglecting the
+    inter-site matrix elements of all Hamiltonians.  This error is
+    discussed in Appendix E, and leads to theoretical uncertainties
+    represented by error bars on the $\O\p{G^3}$ theory in figure 4.
+    A summary of figure 4 is provided in figure 5.  We note that
+    many-body interaction energy shifts are smaller for asymmetric
+    ($-$) states than symmetric ($+$) ones due to the competition
+    between contributions of opposite sign in the asymmetric case (see
+    rows 2 and 3 of table I).}
 
@@
     order for the description of multi-body interactions in the
-    low-energy effective theory.  As experiments begin to work with
-    more atoms and shallower lattices, therefore, reliably predicting
-    interaction energies may require going to yet higher orders in
-    perturbation theory.  Due to a combinatorial explosion of the
-    number of diagrams which appear at increasing orders in the
-    effective theory, however, we need more systematic methods to
-    compute effective multi-body Hamiltonians at fourth order.  In any
-    case, we are agnostic as to whether such a calculation would
-    provide better agreement between experiment and theory without
-    first performing a more detailed analysis of systematic errors.}
+    low-energy effective theory.  The formally small quantities
+    organizing our perturbation theory are two-body ground-state
+    interaction energies (proportional to the couplings $G_X$) divided
+    by the spectral gap of the single-atom Hamiltonian $H_0$.  These
+    reduced (dimensionless) interaction energies vary from
+    $\sim0.05-0.15$ in the parameter regimes of the ${}^{87}$Sr
+    experiments considered here (see Appendix F).  As experiments
+    begin to work with more atoms and different lattice depths,
+    therefore, reliably predicting interaction energies may require
+    going to yet higher orders in perturbation theory.  Due to a
+    combinatorial explosion of the number of diagrams which appear at
+    increasing orders in the effective theory, however, we need more
+    systematic methods to compute effective multi-body Hamiltonians at
+    fourth order.  In any case, we are agnostic as to whether such a
+    calculation would provide better agreement between experiment and
+    theory without first performing a detailed analysis of systematic
+    errors.}
 
-  We hope that this revision has clarified the questions brought up by
-  the referee, and provided a central message for the results in
-  figures 4 and 5.
+  Note that the Appendix F referenced above is not the same as that in
+  the original version of our manuscript, but a new appendix discussed
+  in point \ref{pt:pert_params} below.  We hope that the new text
+  above adequately addresses the questions brought up by the referee,
+  and provides a clear central message for the results in figures 4
+  and 5.
 
@@
 
+  \label{pt:pert_params}
+
+  We thank the referee for pointing out that we had overlooked this
+  important bit of information for interpreting the numerical results
+  in our manuscript.  Consequently, we now provide typical values for
+  our perturbative parameters in the discussion of figures 4 and 5
+  (see point \ref{pt:figs} above), and we have added the following new
+  Appendix to our manuscript:
+
+  \setcounter{figure}{0}
+  \renewcommand\thefigure{F.\arabic{figure}}
+
+  \begin{figure}
+    \centering
+    \includegraphics{../figures/perturbative_params.pdf}
+    \caption{\footnotesize Dependence of the perturbative parameters
+      $KG_X/\Delta$ on the lattice depth $\U$.}
+  \end{figure}
+
+  \green{{\bf Appendix F: Perturbative parameters for the effective
+      theory}}
+
+  \green{The perturbative effective theory developed in Section III is
+    organized in powers of the coupling constants $G_X$.  The formally
+    small, dimensionless quantities for this perturbation theory are
+    the two-body interaction energies $K G_X$, where $K$ is a
+    ground-state two-body overlap integral and $G_X$ is a coupling
+    constant, divided by the spectral gap $\Delta$ of the
+    non-interacting Hamiltonian $H_0$.  The quantities $K$, $G_X$, and
+    $\Delta$ all depend on the lattice depth $\U$.  Figure F.1 shows
+    these parameters for the case of ${}^{87}$Sr atoms with
+    $X\in\set{\g\g, \e\g_-, \e\g_+, \e\e}$ at lattice depths
+    $\U\in\sp{30,80}E_{\t{R}}$, where
+    $E_{\t{R}}\approx3.5\times2\pi~\t{kHz}$ is the lattice photon
+    recoil energy of the atoms.  The fact that these perturbative
+    parameters grow with increasing lattice depth $\U$ is a
+    consequence of the fact that the overlap integral $K$ grows faster
+    with $\U$ than the spectral gap $\Delta$.  In the case of a
+    harmonic trap with angular frequency $\omega$, for example, these
+    parameters would be
+    \begin{multline*}
+      \f{K_{\t{HO}} G_X}{\omega}
+      = \f{G_X}{\omega} \int \d^3 x~\abs{\phi_{\t{HO}}}^4
+      = \f{G_X}{\omega} \sp{\int \d x~
+        \abs{\p{\f{\mA\omega}{\pi}}^{1/4} e^{-\mA\omega x^2/2}}^4}^3 \\
+      \sim \f1{\omega}
+      \sp{\omega \int\f{\d\p{\sqrt\omega x}}{\sqrt\omega}
+        ~\abs{e^{-\mA\omega x^2/2}}^4}^3
+      \propto \sqrt{\omega},
+      \tag{F1}
+    \end{multline*}
+    where we assumed that the coupling constants $G_X$ vary weakly
+    with $\omega$.  While this result may seem to suggest that the
+    low-energy effective theory should become better at smaller
+    lattice depths, smaller lattice depths also result in increased
+    theoretical uncertainties from the growing relevance of the
+    inter-site effects discussed in Appendix E.}
+
+  We hope that these additions help improve readers' interpretations
+  of the numerical results in our manuscript.
+
+\item Finally:
+
+  \blue{Overall, the manuscript contains substantial and solid results
+    and data on few-body spectra of SU(N) atoms and can be quite
+    useful to future studies on these atoms. I recommend the
+    publication of the manuscript in NJP after my concerns above are
+    addressed.}
+
+  We thank the referee for remarking that our manuscript ``contains
+  substantial and results'', and hope that our results will indeed be
+  ``quite useful to future studies''.  We hope that the revisions we
+  outlined above have addressed the referee's concerns and suggestions
+  to a satisfactory degree,
+
 \end{enumerate}

### 2019-02-14 multi_body_interactions/papers/theory/paper.tex
added last of the revisions to the paper (pre-markup by AMR)

@@
 (proportional to the couplings $G_X$) divided by the spectral gap of
-the single-atom Hamiltonian $H_0$.  If $M$ atoms can only couple
-through terms represented by a $p$-vertex diagrams for
-$p\ge p_M^{\t{min}}$, then the leading order contribution to $M$-body
-interactions is order $p_M^{\t{min}}$ in the couplings $G_X$.  If the
-same $p_M^{\t{min}}$-vertex diagrams involve any counter-terms,
-however, then these diagrams are at least order $p_M^{\t{min}}+1$ in
-the couplings $G_X$.  Counter-terms therefore only appear at
+the single-atom Hamiltonian $H_0$ (see Appendix
+\ref{sec:pert_params}).  If $M$ atoms can only couple through terms
+represented by a $p$-vertex diagrams for $p\ge p_M^{\t{min}}$, then
+the leading order contribution to $M$-body interactions is order
+$p_M^{\t{min}}$ in the couplings $G_X$.  If the same
+$p_M^{\t{min}}$-vertex diagrams involve any counter-terms, however,
+then these diagrams are at least order $p_M^{\t{min}}+1$ in the
+couplings $G_X$.  Counter-terms therefore only appear at
 next-to-leading order (NLO) in the calculation of effective $M$-body
@@
   $\Delta_{NX}$ for all $\p{N,X}\in\set{3,4,5}\times\set{+,-}$ at
-  various lattice depths $\U$, and
+  various mean lattice depths $\U$, and
 \item the corresponding values of $\Delta_{NX}$ predicted by the
@@
 description of multi-body interactions in the low-energy effective
-theory.  As experiments begin to work with more atoms and shallower
-lattices, therefore, reliably predicting interaction energies may
+theory.  The formally small quantities organizing our perturbation
+theory are two-body ground-state interaction energies (proportional to
+the couplings $G_X$) divided by the spectral gap of the single-atom
+Hamiltonian $H_0$.  These reduced (dimensionless) interaction energies
+vary from $\sim0.05-0.15$ in the parameter regimes of the ${}^{87}$Sr
+experiments considered here (see Appendix \ref{sec:pert_params}).  As
+experiments begin to work with more atoms and different lattice
+depths, therefore, reliably predicting interaction energies may
 require going to yet higher orders in perturbation theory.  Due to a
@@
 \appendix
+\setcounter{figure}{0}
+\renewcommand\thefigure{\thesection.\arabic{figure}}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Perturbative parameters for the effective theory}
+\label{sec:pert_params}
+
+The perturbative effective theory developed in Section
+\ref{sec:low_energy} is organized in powers of the coupling constants
+$G_X$.  The formally small, dimensionless quantities for this
+perturbation theory are the two-body interaction energies $K G_X$,
+where $K$ is a ground-state two-body overlap integral and $G_X$ is a
+coupling constant, divided by the spectral gap $\Delta$ of the
+non-interacting Hamiltonian $H_0$.  The quantities $K$, $G_X$, and
+$\Delta$ all depend on the lattice depth $\U$.  Figure
+\ref{fig:pert_params} shows these parameters for the case of
+${}^{87}$Sr atoms with $X\in\set{\g\g, \e\g_-, \e\g_+, \e\e}$ at
+lattice depths $\U\in\sp{30,80}E_{\t{R}}$, where
+$E_{\t{R}}\approx3.5\times2\pi~\t{kHz}$ is the lattice photon recoil
+energy of the atoms.  The fact that these perturbative parameters grow
+with increasing lattice depth $\U$ is a consequence of the fact that
+the overlap integral $K$ grows faster with $\U$ than the spectral gap
+$\Delta$.  In the case of a harmonic trap with angular frequency
+$\omega$, for example, these parameters would be
+\begin{multline}
+  \f{K_{\t{HO}} G_X}{\omega}
+  = \f{G_X}{\omega} \int \d^3 x~\abs{\phi_{\t{HO}}}^4
+  = \f{G_X}{\omega} \sp{\int \d x~
+    \abs{\p{\f{\mA\omega}{\pi}}^{1/4} e^{-\mA\omega x^2/2}}^4}^3 \\
+  \sim \f1{\omega}
+  \sp{\omega \int\f{\d\p{\sqrt\omega x}}{\sqrt\omega}
+    ~\abs{e^{-\mA\omega x^2/2}}^4}^3
+  \propto \sqrt{\omega},
+\end{multline}
+where we assumed that the coupling constants $G_X$ vary weakly with
+$\omega$.  While this result may seem to suggest that the low-energy
+effective theory should become better at smaller lattice depths,
+smaller lattice depths also result in increased theoretical
+uncertainties from the growing relevance of the inter-site effects
+discussed in Appendix \ref{sec:error}.
+
+\begin{figure}
+  \centering
+  \includegraphics{perturbative_params.pdf}
+  \caption{\footnotesize Dependence of the perturbative parameters
+    $KG_X/\Delta$ on the lattice depth $\U$.}
+  \label{fig:pert_params}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Low-excitation $M$-body Hamiltonian coefficients}
@@
 
-The $M$-body Hamiltonians addressing low-lying orbital state
-excitations can be written in the form
+When restricted to the subspace of at most one orbital excitation per
+lattice site, the $M$-body Hamiltonians of the low-energy effective
+theory developed in Section \ref{sec:low_energy} can be written in the
+form
 \begin{align}

### 2019-02-14 multi_body_interactions/papers/theory/reviews_NJP/revision_summary.tex
added last of the revisions to the paper (pre-markup by AMR)

@@
     $G_X$) divided by the spectral gap of the single-atom Hamiltonian
-    $H_0$.  If $M$ atoms can only couple through terms represented by
-    a $p$-vertex diagrams for $p\ge p_M^{\t{min}}$, then the leading
-    order contribution to $M$-body interactions is order
-    $p_M^{\t{min}}$ in the couplings $G_X$.  If the same
-    $p_M^{\t{min}}$-vertex diagrams involve any counter-terms,
-    however, then these diagrams are at least order $p_M^{\t{min}}+1$
-    in the couplings $G_X$.  Counter-terms therefore only appear at
-    next-to-leading order (NLO) in the calculation of effective
-    $M$-body interactions.}
+    $H_0$ (see Appendix F).  If $M$ atoms can only couple through
+    terms represented by a $p$-vertex diagrams for
+    $p\ge p_M^{\t{min}}$, then the leading order contribution to
+    $M$-body interactions is order $p_M^{\t{min}}$ in the couplings
+    $G_X$.  If the same $p_M^{\t{min}}$-vertex diagrams involve any
+    counter-terms, however, then these diagrams are at least order
+    $p_M^{\t{min}}+1$ in the couplings $G_X$.  Counter-terms therefore
+    only appear at next-to-leading order (NLO) in the calculation of
+    effective $M$-body interactions.}
 
@@
 
-\item To make our first appendix more self-contained, we have defined
-  the projector $\P_0$ used therein, such that the text (page 30, line
-  38):
-
-  \red{Letting $\Q_0\equiv1-\P_0$ denote the projector onto $\E_0$ and
-    $X$ ...}
-
-  now reads:
-
-  \green{Letting $\P_0$ denote the projector onto $\G_0$,
-    $\Q_0\equiv\1-\P_0$ denote the projector onto $\E_0$, and $X$ ...}
-
 \item For clarity, we have replaced a few instances of the text
@@
 
+\item To make our first appendix more self-contained, we have defined
+  the projector $\P_0$ used therein, such that the text (page 30, line
+  38):
+
+  \red{Letting $\Q_0\equiv1-\P_0$ denote the projector onto $\E_0$ and
+    $X$ ...}
+
+  now reads:
+
+  \green{Letting $\P_0$ denote the projector onto $\G_0$,
+    $\Q_0\equiv\1-\P_0$ denote the projector onto $\E_0$, and $X$ ...}
+
+\item To be more explicit about the restrictions made in Eq.~(F1) in
+  Appendix F, we have changed the text (page 40, line 17):
+
+  \red{The $M$-body Hamiltonians addressing low-lying orbital state
+    excitations can be writtenin the form}
+
+  to:
+
+  \green{When restricted to the subspace of at most one orbital
+    excitation per lattice site, the $M$-body Hamiltonians of the
+    low-energy effective theory developed in Section III can be
+    written in the form}
+
+  Note that due to the addition of a new appendix in the revised
+  version of our manuscript, the above text now appears in Appendix G.
+
 \item Reference 34 of the original manuscript has been published since

### 2019-02-14 multi_body_interactions/papers/theory/reviews_NJP/referee_response_2.tex
added closing remarks to referee 2 (pre-markup by AMR)

@@
 
+  We thank the referee for their consideration
+
   We thank the referee for remarking that our manuscript ``contains
   substantial and results'', and hope that our results will indeed be
-  ``quite useful to future studies''.  We hope that the revisions we
-  outlined above have addressed the referee's concerns and suggestions
-  to a satisfactory degree,
+  ``quite useful to future studies''.  The referee has brought up some
+  fair questions and suggestions, which we have addressed by
+  additional discussion and a new appendix in our manuscript.  We hope
+  that these changes have addressed the referee's concerns to a
+  satisfactory degree, deeming our manuscript suitable for publication
+  in New Journal of Physics.
 

### 2019-02-18 multi_body_interactions/papers/theory/reviews_NJP/referee_response_2.tex
final versions of responses and revisions to NJP

@@
   Specifically, there is a competition between contributions of
-  opposite sign to the interaction energy of asymmetric states.
-  Concerning higher-order effects, it is not obvious when exactly
-  further corrections ($\sim G^4$) should become relevant.  Computing
-  higher-order corrections requires more systematic methods than those
-  used in our manuscript, and we are simply unsure whether considering
-  these corrections is worthwhile without first performing a more
-  detailed analysis of systematic errors in both theory and
+  opposite sign to the interaction energy of asymmetric states.  This
+  competition is a many-body analogue of the two-body case with a
+  competition between direct and exchange terms in the interaction
+  energies of singly-excited states.
+
+  Concerning higher-order effects, it is not obvious to us when
+  exactly further corrections ($\sim G^4$) should become relevant.
+  Computing higher-order corrections requires more systematic methods
+  than those used in our manuscript, and we are simply unsure whether
+  considering these corrections is worthwhile without first performing
+  a more detailed analysis of systematic errors in both theory and
   experiment.
@@
     between contributions of opposite sign in the asymmetric case (see
-    rows 2 and 3 of table I).}
+    rows 2 and 3 of table I, where as a consequence of positive
+    scattering lengths in the case of $^{87}$Sr, all $U_{MX}$ for
+    fixed $M$ have the same sign).  This competition is a many-body
+    analogue of the two-body case with a competition between direct
+    and exchange terms in the interaction energies of singly-excited
+    states.}
 
@@
     experiments considered here (see Appendix F).  As experiments
-    begin to work with more atoms and different lattice depths,
-    therefore, reliably predicting interaction energies may require
-    going to yet higher orders in perturbation theory.  Due to a
-    combinatorial explosion of the number of diagrams which appear at
-    increasing orders in the effective theory, however, we need more
-    systematic methods to compute effective multi-body Hamiltonians at
-    fourth order.  In any case, we are agnostic as to whether such a
-    calculation would provide better agreement between experiment and
-    theory without first performing a detailed analysis of systematic
-    errors.}
+    begin to operate at higher atom densities with amplified
+    interaction effects, reliably predicting interaction energies may
+    require going to yet higher orders in perturbation theory.  Due to
+    a combinatorial explosion of the number of diagrams which appear
+    at increasing orders in the effective theory, however, we need
+    more systematic methods to compute effective multi-body
+    Hamiltonians at fourth order.  In any case, we are agnostic as to
+    whether such a calculation would provide better agreement between
+    experiment and theory without first performing a detailed analysis
+    of systematic errors.}
 
@@
   our perturbative parameters in the discussion of figures 4 and 5
-  (see point \ref{pt:figs} above), and we have added the following new
-  Appendix to our manuscript:
+  (see point \ref{pt:figs} above), where we also direct readers to
+  the following new Appendix in our manuscript:
 
@@
 
-  \begin{figure}
+  \begin{figure}[b]
     \centering
@@
     small, dimensionless quantities for this perturbation theory are
-    the two-body interaction energies $K G_X$, where $K$ is a
-    ground-state two-body overlap integral and $G_X$ is a coupling
-    constant, divided by the spectral gap $\Delta$ of the
-    non-interacting Hamiltonian $H_0$.  The quantities $K$, $G_X$, and
-    $\Delta$ all depend on the lattice depth $\U$.  Figure F.1 shows
-    these parameters for the case of ${}^{87}$Sr atoms with
+    the two-body interaction energies $K G_X$ divided by the spectral
+    gap $\Delta$ of the non-interacting Hamiltonian $H_0$.  Here $K$
+    is a ground-state two-body overlap integral and $G_X$ is a
+    coupling constant.  The quantities $K$, $G_X$, and $\Delta$ all
+    depend on the lattice depth $\U$.  Figure F.1 shows these
+    parameters for the case of ${}^{87}$Sr atoms with
     $X\in\set{\g\g, \e\g_-, \e\g_+, \e\e}$ at lattice depths
@@
     with $\U$ than the spectral gap $\Delta$.  In the case of a
-    harmonic trap with angular frequency $\omega$, for example, these
-    parameters would be
-    \begin{multline*}
+    harmonic trap with angular frequency $\omega$, for example, by
+    dimensional analysis these parameters would be
+    \begin{align*}
       \f{K_{\t{HO}} G_X}{\omega}
@@
       = \f{G_X}{\omega} \sp{\int \d x~
-        \abs{\p{\f{\mA\omega}{\pi}}^{1/4} e^{-\mA\omega x^2/2}}^4}^3 \\
-      \sim \f1{\omega}
-      \sp{\omega \int\f{\d\p{\sqrt\omega x}}{\sqrt\omega}
-        ~\abs{e^{-\mA\omega x^2/2}}^4}^3
-      \propto \sqrt{\omega},
+        \abs{\p{\f{\mA\omega}{\pi}}^{1/4} e^{-\mA\omega x^2/2}}^4}^3
+      \sim \omega^{1/2},
       \tag{F1}
-    \end{multline*}
+    \end{align*}
     where we assumed that the coupling constants $G_X$ vary weakly
@@
 
-  We thank the referee for their consideration
-
   We thank the referee for remarking that our manuscript ``contains
-  substantial and results'', and hope that our results will indeed be
-  ``quite useful to future studies''.  The referee has brought up some
-  fair questions and suggestions, which we have addressed by
-  additional discussion and a new appendix in our manuscript.  We hope
-  that these changes have addressed the referee's concerns to a
-  satisfactory degree, deeming our manuscript suitable for publication
-  in New Journal of Physics.
+  substantial and solid results'', and hope that our results will
+  indeed be ``quite useful to future studies''.  The referee has
+  brought up several good questions and suggestions, which we have
+  addressed through additional discussion and a new appendix in our
+  manuscript.  We hope that these changes have addressed the referee's
+  concerns to a satisfactory degree, deeming our manuscript suitable
+  for publication in New Journal of Physics.
 

### 2019-02-18 multi_body_interactions/papers/theory/paper.tex
final versions of responses and revisions to NJP

@@
     given in terms of the coefficients $U_{MX}$ as appearing in
-    \eqref{eq:H_M}, in addition to the $M$-body eigenvalues
-    $E_{MX}^{(M)}$.}
+    \eqref{eq:H_M} (first three rows), in addition to the $M$-body
+    eigenvalues $E_{MX}^{(M)}$ (last three rows).}
   \label{tab:eigen}
@@
     Eigenstate
-    & $H_M$ eigenvalue \\ \hline\hline
+    & $H_M$ eigenvalue ($M\le N$) \\ \hline\hline
     $\ket{\N,0}$
@@
 opposite sign in the asymmetric case (see rows 2 and 3 of table
-\ref{tab:eigen}).
+\ref{tab:eigen}, where as a consequence of positive scattering lengths
+in the case of $^{87}$Sr, all $U_{MX}$ for fixed $M$ have the same
+sign).  This competition is a many-body analogue of the two-body case
+with a competition between direct and exchange terms in the
+interaction energies of singly-excited states.
 
@@
 experiments considered here (see Appendix \ref{sec:pert_params}).  As
-experiments begin to work with more atoms and different lattice
-depths, therefore, reliably predicting interaction energies may
+experiments begin to operate at higher atom densities with amplified
+interaction effects, reliably predicting interaction energies may
 require going to yet higher orders in perturbation theory.  Due to a
@@
 $G_X$.  The formally small, dimensionless quantities for this
-perturbation theory are the two-body interaction energies $K G_X$,
-where $K$ is a ground-state two-body overlap integral and $G_X$ is a
-coupling constant, divided by the spectral gap $\Delta$ of the
-non-interacting Hamiltonian $H_0$.  The quantities $K$, $G_X$, and
-$\Delta$ all depend on the lattice depth $\U$.  Figure
+perturbation theory are the two-body interaction energies $K G_X$
+divided by the spectral gap $\Delta$ of the non-interacting
+Hamiltonian $H_0$.  Here $K$ is a ground-state two-body overlap
+integral and $G_X$ is a coupling constant.  The quantities $K$, $G_X$,
+and $\Delta$ all depend on the lattice depth $\U$.  Figure
 \ref{fig:pert_params} shows these parameters for the case of
@@
 $\Delta$.  In the case of a harmonic trap with angular frequency
-$\omega$, for example, these parameters would be
-\begin{multline}
+$\omega$, for example, by dimensional analysis these parameters would
+be
+\begin{align}
   \f{K_{\t{HO}} G_X}{\omega}
@@
   = \f{G_X}{\omega} \sp{\int \d x~
-    \abs{\p{\f{\mA\omega}{\pi}}^{1/4} e^{-\mA\omega x^2/2}}^4}^3 \\
-  \sim \f1{\omega}
-  \sp{\omega \int\f{\d\p{\sqrt\omega x}}{\sqrt\omega}
-    ~\abs{e^{-\mA\omega x^2/2}}^4}^3
+    \abs{\p{\f{\mA\omega}{\pi}}^{1/4} e^{-\mA\omega x^2/2}}^4}^3
   \propto \sqrt{\omega},
-\end{multline}
+\end{align}
 where we assumed that the coupling constants $G_X$ vary weakly with
@@
   U_{4,\g}^{(3)}
-  = \p{\alpha_{4,1}^{(3)} - \alpha_5^{(3)}} G_\g^3
+  = \p{2\alpha_{4,1}^{(3)} - \alpha_5^{(3)}} G_\g^3
   + \p{\alpha_{4,2}^{(3)} - \alpha_5^{(3)}} 2 G_\g^3,
@@
   U_{4,+}^{(3)}
-  &= \p{\alpha_{4,1}^{(3)} - \alpha_5^{(3)}}
+  &= \p{2\alpha_{4,1}^{(3)} - \alpha_5^{(3)}}
   2 G_+ G_\g \p{G_+ + G_\g} \nonumber \\
@@
   U_{4,-}^{(3)}
-  &= \p{\alpha_{4,1}^{(3)} - \alpha_5^{(3)}}
+  &= \p{2\alpha_{4,1}^{(3)} - \alpha_5^{(3)}}
   2 G_- G_\g \p{2 G_+ + G_- + G_\g} \nonumber \\

### 2019-02-18 multi_body_interactions/papers/theory/reviews_NJP/revision_summary.tex
final versions of responses and revisions to NJP

@@
 \begin{enumerate}[label=(R2.\arabic{point}.\arabic*)]
-\item
+  \step
+\item We have made the caption of figure 1 more self-contained,
+  replacing the text (page 11, line 13):
+
+  \red{... $n>0$ and $\c_{\mu s}\equiv\c_{0,\mu s}$.  For the sake
+    presentation, ...}
+
+  by:
+
+  \green{... $n>0$ and $\c_{\mu s}\equiv\c_{0,\mu s}$.  Diagrams are
+    read from left to right to construct a sequence of operators from
+    right to left.  Solid (dashed) lines represent field operators
+    acting on the lowest (arbitrary) motional states.  For the sake
+    presentation, ...}
+
+  \step
+\item We have rewritten our discussion of figures 4 and 5, replacing
+  the text (page 23, line 54):
+
+  \red{Identifying peaks in excitation spectra such as in figure 3
+    constitutes a measurement of many-body excitation energies, which
+    was performed in ref.~[34] to detect signatures of effective
+    multi-body interactions.  Figure 4 shows a comparison between
+    experimental measurements of the many-body excitation energies
+    $\Delta_{NX}$ for $N\in\set{3,4,5}$ at various lattice depths
+    $\U$, and the corresponding values of $\Delta_{NX}$ predicted by
+    the low-energy effective theory at different orders in the
+    coupling constants.  A summary of the data in figure 4 is provided
+    in figure 5.  These figures show a clear improvement of the
+    low-energy effective theory we have developed with increasing
+    order in the coupling constants, in particular demonstrating the
+    need to consider multi-body interactions in a single-band model of
+    the atoms without motional excitations.}
+
+  \red{We expect that the dominant source of error in the third-order
+    effective theory comes from the fact that we neglect the
+    inter-site matrix elements of all Hamiltonians.  This source of
+    error is discussed in Appendix E, and leads to theoretical
+    uncertainties represented by error bars in figure 4.  As these
+    uncertainties are sometimes unable to account for the residual
+    disagreement between experimental measurements and the effective
+    theory, better agreement requires either a better understanding of
+    experimental procedures and error, a better understanding of
+    theoretical uncertainties, or consideration of yet higher-order
+    terms in the effective theory.  Due to a combinatorial explosion
+    in the number of diagrams which appear at increasing perturbative
+    orders of the effective theory, we need more systematic methods to
+    compute effective multi-body Hamiltonians at fourth order; we
+    therefore leave this calculation for future work.}
+
+  by:
+
+  \green{Identifying peaks in excitation spectra such as in figure 3
+    constitutes a measurement of many-body excitation energies, which
+    was performed in ref.~[34] to detect signatures of effective
+    multi-body interactions.  Figure 4 shows a comparison between (i)
+    experimental measurements of the many-body excitation energies
+    $\Delta_{NX}$ for all $\p{N,X}\in\set{3,4,5}\times\set{+,-}$ at
+    various mean lattice depths $\U$, and (ii) the corresponding
+    values of $\Delta_{NX}$ predicted by the low-energy effective
+    theory at different orders in the coupling constants.  A known
+    source of error in our effective theory comes from neglecting the
+    inter-site matrix elements of all Hamiltonians.  This error is
+    discussed in Appendix E, and leads to theoretical uncertainties
+    represented by error bars on the $\O\p{G^3}$ theory in figure 4.
+    A summary of figure 4 is provided in figure 5.  We note that
+    many-body interaction energy shifts are smaller for asymmetric
+    ($-$) states than symmetric ($+$) ones due to the competition
+    between contributions of opposite sign in the asymmetric case (see
+    rows 2 and 3 of table I, where as a consequence of positive
+    scattering lengths in the case of $^{87}$Sr, all $U_{MX}$ for
+    fixed $M$ have the same sign).  This competition is a many-body
+    analogue of the two-body case with a competition between direct
+    and exchange terms in the interaction energies of singly-excited
+    states.}
+
+  \green{The results in figures 4 and 5 highlight a few important
+    points about ultracold, high-density ${}^{87}$Sr experiments and
+    our low-energy effective theory.  First, these experiments exhibit
+    clear signatures of multi-body interactions, as evidenced by a
+    stark disagreement between the observed many-body excitation
+    energies $\Delta_{NX}$ and those that are predicted by the
+    two-body $\O\p{G}$ theory.  Multi-body interactions are thus
+    crucial for understanding these high-density experiments in the
+    context of a single-band Hubbard model, which naturally arises in
+    the zero-temperature limit when all atoms occupy their motional
+    ground state.  Second, the inter-atomic interactions in these
+    experiments are strong enough to require going beyond leading
+    order for the description of multi-body interactions in the
+    low-energy effective theory.  The formally small quantities
+    organizing our perturbation theory are two-body ground-state
+    interaction energies (proportional to the couplings $G_X$) divided
+    by the spectral gap of the single-atom Hamiltonian $H_0$.  These
+    reduced (dimensionless) interaction energies vary from
+    $\sim0.05-0.15$ in the parameter regimes of the ${}^{87}$Sr
+    experiments considered here (see Appendix F).  As experiments
+    begin to operate at higher atom densities with amplified
+    interaction effects, reliably predicting interaction energies may
+    require going to yet higher orders in perturbation theory.  Due to
+    a combinatorial explosion of the number of diagrams which appear
+    at increasing orders in the effective theory, however, we need
+    more systematic methods to compute effective multi-body
+    Hamiltonians at fourth order.  In any case, we are agnostic as to
+    whether such a calculation would provide better agreement between
+    experiment and theory without first performing a detailed analysis
+    of systematic errors.}
+
+  Note that the Appendix F referenced above is not the same as that in
+  the original version of our manuscript, but a new appendix that we
+  have added in the revised version.
+
+  \step
+\item We have added the following appendix to our manuscript:
+
+  \setcounter{figure}{0}
+  \renewcommand\thefigure{F.\arabic{figure}}
+
+  \begin{figure}
+    \centering
+    \includegraphics{../figures/perturbative_params.pdf}
+    \caption{\footnotesize Dependence of the perturbative parameters
+      $KG_X/\Delta$ on the lattice depth $\U$.}
+  \end{figure}
+
+  \green{{\bf Appendix F: Perturbative parameters for the effective
+      theory}}
+
+  \green{The perturbative effective theory developed in Section III is
+    organized in powers of the coupling constants $G_X$.  The formally
+    small, dimensionless quantities for this perturbation theory are
+    the two-body interaction energies $K G_X$ divided by the spectral
+    gap $\Delta$ of the non-interacting Hamiltonian $H_0$.  Here $K$
+    is a ground-state two-body overlap integral and $G_X$ is a
+    coupling constant.  The quantities $K$, $G_X$, and $\Delta$ all
+    depend on the lattice depth $\U$.  Figure F.1 shows these
+    parameters for the case of ${}^{87}$Sr atoms with
+    $X\in\set{\g\g, \e\g_-, \e\g_+, \e\e}$ at lattice depths
+    $\U\in\sp{30,80}E_{\t{R}}$, where
+    $E_{\t{R}}\approx3.5\times2\pi~\t{kHz}$ is the lattice photon
+    recoil energy of the atoms.  The fact that these perturbative
+    parameters grow with increasing lattice depth $\U$ is a
+    consequence of the fact that the overlap integral $K$ grows faster
+    with $\U$ than the spectral gap $\Delta$.  In the case of a
+    harmonic trap with angular frequency $\omega$, for example, by
+    dimensional analysis these parameters would be
+    \begin{align*}
+      \f{K_{\t{HO}} G_X}{\omega}
+      = \f{G_X}{\omega} \int \d^3 x~\abs{\phi_{\t{HO}}}^4
+      = \f{G_X}{\omega} \sp{\int \d x~
+        \abs{\p{\f{\mA\omega}{\pi}}^{1/4} e^{-\mA\omega x^2/2}}^4}^3
+      \sim \omega^{1/2},
+      \tag{F1}
+    \end{align*}
+    where we assumed that the coupling constants $G_X$ vary weakly
+    with $\omega$.  While this result may seem to suggest that the
+    low-energy effective theory should become better at smaller
+    lattice depths, smaller lattice depths also result in increased
+    theoretical uncertainties from the growing relevance of the
+    inter-site effects discussed in Appendix E.}
+
 \end{enumerate}
@@
 
+\item For clarity, we have replaced a few instances of the text
+  ``\red{single-particle}'' or ``\red{single-atom}'' by
+  ``\green{non-interacting}'' when referring to motional ground states
+  of non-interacting particles; specifically, we have done so on page
+  9, line 50; on page 11, line 22; on page 13, line 5; and on page 35,
+  line 26.
+
 \item The last two lines of page 10 and line 26 of page 11 previously
@@
   ordering convention is used throughout the rest of the manuscript:
-  $G^{\mu q;\nu r}_{\rho s;\sigma t}$ is gives the coupling constant
-  for destroying two atoms in state $\p{\mu q}+\p{\nu r}$, and
-  creating them in $\p{\rho s}+\p{\sigma t}$.
+  $G^{\mu q;\nu r}_{\rho s;\sigma t}$ gives the coupling constant for
+  destroying two atoms in state $\p{\mu q}+\p{\nu r}$, and creating
+  them in $\p{\rho s}+\p{\sigma t}$.
 
@@
 
+\item In table I, we have change the following sentence of the caption
+  (page 20, line 7):
+
+  \red{The corresponding $N$-body eigenvalues $E_{NX}^{(M)}$ of $H_M$
+    are given in terms of the coefficients $U_{MX}$ as appearing in
+    (42), in addition to the $M$-body eigenvalues $E_{MX}^{(M)}$.}
+
+  to:
+
+  \green{The corresponding $N$-body eigenvalues $E_{NX}^{(M)}$ of
+    $H_M$ are given in terms of the coefficients $U_{MX}$ as appearing
+    in (44) (first three rows), in addition to the $M$-body
+    eigenvalues $E_{MX}^{(M)}$ (last three rows).}
+
+  Note that the equation referenced in this text has only changed due
+  to the addition of two more equations in the revised version of our
+  manuscript.  In addition, we have made explicit that this table
+  applies for $M\le N$, by changing the column header (page 20, line
+  13) ``\red{$H_M$ eigenvalue}'' to ``\green{$H_M$ eigenvalue
+    ($M\le N$)}''.
+
 \item In figure 2 (page 22), all multiplicities (i.e.~values on the
@@
 
-\item For clarity, we have replaced a few instances of the text
-  ``\red{single-particle}'' or ``\red{single-atom}'' by
-  ``\green{non-interacting}'' when referring to motional ground states
-  of non-interacting particles; specifically, we have done so on page
-  9, line 50; on page 11, line 22; on page 13, line 5; and on page 35,
-  line 26.
-
 \item To make our first appendix more self-contained, we have defined
@@
 
+\item We have fixed an error in our list of low-excitation $M$-body
+  Hamiltonian coefficients, replacing
+  \red{$\p{\alpha_{4,1}^{(3)} - \alpha_5^{(3)}}$} by
+  \green{$\p{2\alpha_{4,1}^{(3)} - \alpha_5^{(3)}}$} in Eqs.~(F8),
+  (F9), and (F10) on page 41.  Note that these equations are
+  respectively (G8), (G9), and (G10) in the revised manuscript.
+
 \item Reference 34 of the original manuscript has been published since

### 2019-02-27 multi_body_interactions/papers/theory/paper.tex
fixed error in multi-body thoery error concerned incorrect renormalization of third-order three-body loop diagrams

@@
 renormalize our coupling constants by introducing counter-terms
-$\tilde G_X$ into the interaction Hamiltonian.
+$\widetilde G_X$ into the interaction Hamiltonian.
 
@@
 the ``bare'' coupling constants $G_X^{\t{bare}}$ that are used in
-\eqref{eq:H_2_expansion} as $G_X^{\t{bare}} = G_X + \tilde G_X$.  In
-performing such a decomposition, we are free to choose the values of
-$G_X$, which in turn fixes the values of
-$\tilde G_X \equiv G_X^{\t{bare}} - G_X$.  For convenience, we can
+\eqref{eq:H_2_expansion} as $G_X^{\t{bare}} = G_X + \widetilde G_X$.
+In performing such a decomposition, we are free to choose the values
+of $G_X$, which in turn fixes the values of
+$\widetilde G_X \equiv G_X^{\t{bare}} - G_X$.  For convenience, we can
 choose the values of $G_X$ to be those of the net effective coupling
@@
 Representing the new coupling constants $G_X$ by regular vertices and
-the counter-terms $\tilde G_X$ by a crossed dot (i.e.~$\otimes$), this
-choice leads to the {\it renormalization condition}
+the counter-terms $\widetilde G_X$ by a crossed dot (i.e.~$\otimes$),
+this choice leads to the {\it renormalization condition}
 \begin{align}
@@
       \end{tikzpicture}}
-    \\ \O(\tilde G)}
+    \\ \O(\widetilde G)}
   - \substack{
@@
 Second, the renormalization condition in \eqref{eq:counter_term}
-implies that the counter terms $\tilde G_X$ are second order in the
-coupling constants $G_X$, i.e.~$\tilde G_X\sim G_X^2$.  Although the
-effective Hamiltonian expansions in \eqref{eq:H_int_1_2} and
-\eqref{eq:H_int_3} are organized in powers of the interaction
+implies that the counter terms $\widetilde G_X$ are second order in
+the coupling constants $G_X$, i.e.~$\widetilde G_X\sim G_X^2$.
+Although the effective Hamiltonian expansions in \eqref{eq:H_int_1_2}
+and \eqref{eq:H_int_3} are organized in powers of the interaction
 Hamiltonian $H_{\t{int}}$, the couplings $G_X$ are the ``small''
@@
   \end{tikzpicture}
-  \propto K_n^2 \tilde G^{rs}_{r's'} G^{s't}_{s''t'}
+  \propto K_n^2 \widetilde G^{rs}_{r's'} G^{s't}_{s''t'}
   \c_{\mu r'}^\dag \c_{\nu s''}^\dag \c_{\rho t'}^\dag
@@
 \end{align}
-and its mirror image, where $\tilde G^{qr}_{st}$ is equal to the
-counter-term associated with $G^{qr}_{st}$.  To leading order in the
-coupling constants, the renormalization condition in
-\eqref{eq:counter_term} implies that
+and its mirror image, where $\widetilde G^{qr}_{st}$ is equal to the
+counter-term associated with $G^{qr}_{st}$.
+
+The net contribution to the third-order three-body interaction
+Hamiltonian from three-particle-loop diagrams of the form in
+\eqref{eq:H_3_3_S} is
+\begin{align}
+  \shrink{
+    \begin{tikzpicture}
+      \begin{feynman}
+        \vertex (v1);
+        \vertex[below right = of v1] (v2);
+        \vertex[above right = of v2] (v3);
+        \vertex[above left = of v1] (f1);
+        \vertex[left = of v1] (f2);
+        \vertex[below left = of v2] (f3);
+        \vertex[above right = of v3] (f4);
+        \vertex[right = of v3] (f5);
+        \vertex[below right = of v2] (f6);
+        \diagram* {
+          (f1) -- (v1) --[scalar] (v3) -- (f4),
+          (f2) -- (v1) --[scalar] (v2) --[scalar] (v3) -- (f5),
+          (f3) -- (v2) -- (f6), };
+      \end{feynman}
+    \end{tikzpicture}}
+  - \f12 \shrink{
+    \begin{tikzpicture}
+      \begin{feynman}
+        \vertex (v1);
+        \vertex[below right = of v1] (v2);
+        \vertex[above right = of v2] (v3);
+        \vertex[above left = of v1] (f1);
+        \vertex[left = of v1] (f2);
+        \vertex[below left = of v2] (f3);
+        \vertex[above right = of v3] (f4);
+        \vertex[right = of v3] (f5);
+        \vertex[below right = of v2] (f6);
+        \diagram* {
+          (f1) -- (v1) -- (v3) -- (f4),
+          (f2) -- (v1) --[scalar] (v2) -- (v3) -- (f5),
+          (f3) -- (v2) -- (f6), };
+      \end{feynman}
+    \end{tikzpicture}}
+  - \f12 \shrink{
+    \begin{tikzpicture}
+      \begin{feynman}
+        \vertex (v1);
+        \vertex[below right = of v1] (v2);
+        \vertex[above right = of v2] (v3);
+        \vertex[above left = of v1] (f1);
+        \vertex[left = of v1] (f2);
+        \vertex[below left = of v2] (f3);
+        \vertex[above right = of v3] (f4);
+        \vertex[right = of v3] (f5);
+        \vertex[below right = of v2] (f6);
+        \diagram* {
+          (f1) -- (v1) -- (v3) -- (f4),
+          (f2) -- (v1) -- (v2) --[scalar] (v3) -- (f5),
+          (f3) -- (v2) -- (f6), };
+      \end{feynman}
+    \end{tikzpicture}}
+  = \p{\alpha_{3,1}^{(3)} - \alpha_5^{(3)}} \H_{3,1}^{(3)},
+\end{align}
+where
+\begin{align}
+  \alpha_{3,1}^{(3)} \equiv \sum_{\substack{\ell+m>0\\\ell+n>0}}
+  \f{K_{\ell m} K^m_n K_{\ell n}}{E_{\ell m} E_{\ell n}},
+  &&
+  \alpha_5^{(3)}
+  \equiv  K \sum_{n>0} \f{K_n^2}{E_n^2},
+  \label{eq:alph_3_3_1}
+\end{align}
+and
+\begin{align}
+  \H_{3,1}^{(3)} \equiv \sum_{\abs{\set{\mu,\nu,\rho}}=3}
+  G^{rs}_{r's'} G^{\nu s'\rho t}_{\nu's''\rho't'} G^{r's''}_{r''s'''}
+  \c_{\mu r''}^\dag \c_{\nu's'''}^\dag \c_{\rho't'}^\dag
+  \c_{\rho t} \c_{\nu s} \c_{\mu r}.
+\end{align}
+Even though this contribution comes from loop diagrams, the factors
+$\alpha_{3,1}^{(3)}$ and $\alpha_5^{(3)}$ in \eqref{eq:alph_3_3_1} are
+finite.  At large motional state indices $n$, atoms become free
+particles for which $n$ essentially indexes discrete momentum states.
+These atoms thus have an energy which asymptotically scales as
+$E_n\sim n^2\equiv n_\x^2+n_\y^2+n_\z^2$.  Furthermore, the
+oscillatory behavior of atomic wavefunctions with increasing motional
+state indices $\ell,m$ implies that the overlap integral $K_{\ell m}$
+becomes sharply peaked at $\ell\approx m$ as $\ell$ and $m$ get large.
+The asymptotic behavior of $\alpha_{3,1}^{(3)}$ at large $\ell,m,n$ is
+therefore
+\begin{align}
+  \alpha_{3,1}^{(3)}
+  \sim \int \f{\d^3\ell~\d^3m~\d^3n}{\p{\ell^2+m^2}\p{\ell^2+n^2}}
+  ~\delta\p{\ell-m}\delta\p{\ell-n}
+  \sim \int \f{\d^3\ell}{\ell^4}
+  \sim \int_{\ell_{\t{min}}}^\infty \f{\d\ell}{\ell^2}
+  \sim \f1{\ell_{\t{min}}},
+  \label{eq:alph_3_3_1_integral}
+\end{align}
+where in the last integral we changed to spherical coordinates, and
+$\ell_{\t{min}}^2$ is the minimum value of $\ell^2$ for which
+\begin{enumerate*}
+\item the energy $E_\ell\sim\ell^2$, and
+\item the integral expression in \eqref{eq:alph_3_3_1_integral} is a
+  good approximation to the corresponding sum in
+  \eqref{eq:alph_3_3_1}.
+\end{enumerate*}
+Note that the introduction of $\ell_{\t{min}}$ amounts to neglecting a
+finite number of terms in the sum over $\ell,m,n$ in
+\eqref{eq:alph_3_3_1}, whose contribution to the value of
+$\alpha_{3,1}^{(3)}$ is finite.  Convergence of $\alpha_5^{(3)}$ is
+similarly guaranteed by the fact that the overlap integral $K_n$ does
+not asymptotically grow with increasing $n$, such that
+$\alpha_5^{(3)}$ asymptotically behaves as
+\begin{align}
+  \alpha_5^{(3)} \sim \int \f{\d^3 n}{n^4}
+  \sim \int_{n_{\t{min}}}^\infty \f{\d n}{n^2}
+  \sim \f1{n_{\t{min}}},
+  \label{eq:alph_5_3_integral}
+\end{align}
+where again $n_{\t{min}}$ is defined similarly to $\ell_{\t{min}}$.
+
+The sum over loop diagrams in \eqref{eq:H_3_3_O}, meanwhile, contains
+a divergence that must be cancelled out by the counter-terms in
+\eqref{eq:H_3_3_counter_term}.  To leading order in the coupling
+constants, the renormalization condition in \eqref{eq:counter_term}
+implies that
 \begin{align}
@@
 \end{align}
-which means that the sum over all counter-term diagrams in
-\eqref{eq:H_3_3_counter_term} is
-\begin{align}
+which can be expanded and solved explicitly for the counter-terms to
+find
+\begin{align}
+  \widetilde G^{rs}_{r''s''}
+  = \sum_{\ell,m,r',s'} \f{K_{\ell m}^2}{K E_{\ell m}}
+  G^{rs}_{r's'} G^{r's'}_{r''s''}.
+\end{align}
+In terms of ordinary coupling constants, the counter-term diagram in
+\eqref{eq:H_3_3_counter_term} is therefore
+\begin{align}
+  \begin{tikzpicture}
+    \begin{feynman}
+      \vertex[crossed dot] (v1) {};
+      \vertex[above left = 4em of v1] (f1) {$\mu r$};
+      \vertex[below left = 4em of v1] (f2) {$\nu s$};
+      \vertex[right = 4em of v1] (v2);
+      \vertex[above right = 4em of v1] (f3) {$\mu r''$};
+      \vertex[below left = of v2] (f4) {$\rho t$};
+      \vertex[below right = of v2] (f5) {$\rho t'$};
+      \vertex[above right = of v2] (f6) {$\nu s'''$};
+      \diagram* {
+        (f1) -- (v1) -- (f3),
+        (f2) -- (v1)
+        --[scalar, edge label=$n\nu s''$] (v2),
+        (f4) -- (v2) -- (f5),
+        (v2) -- (f6), };
+    \end{feynman}
+  \end{tikzpicture}
+  = \sum_{\ell,m,r',s'}
+  \f{K_{\ell m}^2 K_n^2}{K E_{\ell m} E_n}
+  G^{rs}_{r's'} G^{r's'}_{r''s''} G^{s''t}_{s'''t'}
+  \c_{\mu r''}^\dag \c_{\nu s'''}^\dag \c_{\rho t'}^\dag
+  \c_{\rho t} \c_{\nu s} \c_{\mu r}.
+\end{align}
+Altogether, the contribution to the third-order three-body interaction
+Hamiltonian from loop diagrams of the form in \eqref{eq:H_3_3_O} and
+counter-term diagrams of the form in \eqref{eq:H_3_3_counter_term} is
+\begin{multline}
   \shrink{
+    \begin{tikzpicture}
+      \begin{feynman}
+        \vertex (v1);
+        \vertex[right = of v1] (v2);
+        \vertex[below right = of v2] (v3);
+        \vertex[above left = of v1] (f1);
+        \vertex[below left = of v1] (f2);
+        \vertex[below left = of v3] (f3);
+        \vertex[above right = of v2] (f4);
+        \vertex[above right = of v3] (f5);
+        \vertex[below right = of v3] (f6);
+        \diagram* {
+          (f1) -- (v1) --[half left, scalar] (v2) -- (f4),
+          (f2) -- (v1) --[half right, scalar] (v2)
+          --[scalar] (v3) -- (f5),
+          (f3) -- (v3) -- (f6), };
+      \end{feynman}
+    \end{tikzpicture}}
+    - \shrink{
     \begin{tikzpicture}
@@
     \end{tikzpicture}}
-  = \shrink{
+  - \f12 \shrink{
     \begin{tikzpicture}
@@
           (f2) -- (v1) --[half right, scalar] (v2)
-          --[scalar] (v3) -- (f5),
+          -- (v3) -- (f5),
           (f3) -- (v3) -- (f6), };
       \end{feynman}
-    \end{tikzpicture}}.
-  \label{eq:H_3_3_counter_term_sum}
-\end{align}
-The net contribution of the these counter-term diagrams together with
-the two-particle-loop diagrams in \eqref{eq:H_3_3_O} is therefore
-\begin{multline}
-  - \shrink{
-    \begin{tikzpicture}
-      \begin{feynman}
-        \vertex[crossed dot] (v1) {};
-        \vertex[right = 4em of v1] (v2);
-        \vertex[above left = of v1] (f1);
-        \vertex[below left = of v1] (f2);
-        \vertex[above right = of v1] (f3);
-        \vertex[below left = of v2] (f4);
-        \vertex[below right = of v2] (f5);
-        \vertex[above right = of v2] (f6);
-        \diagram* {
-          (f1) -- (v1) -- (f3),
-          (f2) -- (v1) --[scalar] (v2),
-          (f4) -- (v2) -- (f5),
-          (v2) -- (f6), };
-      \end{feynman}
     \end{tikzpicture}}
-  + \shrink{
+  - \f12 \shrink{
     \begin{tikzpicture}
@@
         \diagram* {
-          (f1) -- (v1) --[half left, scalar] (v2) -- (f4),
-          (f2) -- (v1) --[half right, scalar] (v2)
-          --[scalar] (v3) -- (f5),
-          (f3) -- (v3) -- (f6), };
-      \end{feynman}
-    \end{tikzpicture}}
-  - \f12 \shrink{
-    \begin{tikzpicture}
-      \begin{feynman}
-        \vertex (v1);
-        \vertex[right = of v1] (v2);
-        \vertex[below right = of v2] (v3);
-        \vertex[above left = of v1] (f1);
-        \vertex[below left = of v1] (f2);
-        \vertex[below left = of v3] (f3);
-        \vertex[above right = of v2] (f4);
-        \vertex[above right = of v3] (f5);
-        \vertex[below right = of v3] (f6);
-        \diagram* {
-          (f1) -- (v1) --[half left, scalar] (v2) -- (f4),
-          (f2) -- (v1) --[half right, scalar] (v2)
-          -- (v3) -- (f5),
-          (f3) -- (v3) -- (f6), };
-      \end{feynman}
-    \end{tikzpicture}}
-  - \f12 \shrink{
-    \begin{tikzpicture}
-      \begin{feynman}
-        \vertex (v1);
-        \vertex[right = of v1] (v2);
-        \vertex[below right = of v2] (v3);
-        \vertex[above left = of v1] (f1);
-        \vertex[below left = of v1] (f2);
-        \vertex[below left = of v3] (f3);
-        \vertex[above right = of v2] (f4);
-        \vertex[above right = of v3] (f5);
-        \vertex[below right = of v3] (f6);
-        \diagram* {
           (f1) -- (v1) --[half left] (v2) -- (f4),
@@
     \end{tikzpicture}} \\
-  = - \f12 \shrink{
-    \begin{tikzpicture}
-      \begin{feynman}
-        \vertex (v1);
-        \vertex[right = of v1] (v2);
-        \vertex[below right = of v2] (v3);
-        \vertex[above left = of v1] (f1);
-        \vertex[below left = of v1] (f2);
-        \vertex[below left = of v3] (f3);
-        \vertex[above right = of v2] (f4);
-        \vertex[above right = of v3] (f5);
-        \vertex[below right = of v3] (f6);
-        \diagram* {
-          (f1) -- (v1) --[half left, scalar] (v2) -- (f4),
-          (f2) -- (v1) --[half right, scalar] (v2)
-          -- (v3) -- (f5),
-          (f3) -- (v3) -- (f6), };
-      \end{feynman}
-    \end{tikzpicture}}
-  - \f12 \shrink{
-    \begin{tikzpicture}
-      \begin{feynman}
-        \vertex (v1);
-        \vertex[right = of v1] (v2);
-        \vertex[below right = of v2] (v3);
-        \vertex[above left = of v1] (f1);
-        \vertex[below left = of v1] (f2);
-        \vertex[below left = of v3] (f3);
-        \vertex[above right = of v2] (f4);
-        \vertex[above right = of v3] (f5);
-        \vertex[below right = of v3] (f6);
-        \diagram* {
-          (f1) -- (v1) --[half left] (v2) -- (f4),
-          (f2) -- (v1) --[half right] (v2)
-          --[scalar] (v3) -- (f5),
-          (f3) -- (v3) -- (f6), };
-      \end{feynman}
-    \end{tikzpicture}}.
-  \label{eq:H_3_3_cancellation}
+  = \p{\alpha_{3,2}^{(3)} - \f12\alpha_{4,3}^{(3)} - \f12\alpha_5^{(3)}}
+  \H_{3,2}^{(3)},
 \end{multline}
-An identical cancellation occurs for the mirror images of the diagrams
-in \eqref{eq:H_3_3_O} and \eqref{eq:H_3_3_counter_term}, which results
-in a contribution to the effective Hamiltonian that is equal to
-\eqref{eq:H_3_3_cancellation}.  The contribution from the
-three-particle-loop diagrams in \eqref{eq:H_3_3_S} is
-\begin{multline}
-  \shrink{
-    \begin{tikzpicture}
-      \begin{feynman}
-        \vertex (v1);
-        \vertex[below right = of v1] (v2);
-        \vertex[above right = of v2] (v3);
-        \vertex[above left = of v1] (f1);
-        \vertex[left = of v1] (f2);
-        \vertex[below left = of v2] (f3);
-        \vertex[above right = of v3] (f4);
-        \vertex[right = of v3] (f5);
-        \vertex[below right = of v2] (f6);
-        \diagram* {
-          (f1) -- (v1) --[scalar] (v3) -- (f4),
-          (f2) -- (v1) --[scalar] (v2) --[scalar] (v3) -- (f5),
-          (f3) -- (v2) -- (f6), };
-      \end{feynman}
-    \end{tikzpicture}}
-  - \f12 \shrink{
-    \begin{tikzpicture}
-      \begin{feynman}
-        \vertex (v1);
-        \vertex[below right = of v1] (v2);
-        \vertex[above right = of v2] (v3);
-        \vertex[above left = of v1] (f1);
-        \vertex[left = of v1] (f2);
-        \vertex[below left = of v2] (f3);
-        \vertex[above right = of v3] (f4);
-        \vertex[right = of v3] (f5);
-        \vertex[below right = of v2] (f6);
-        \diagram* {
-          (f1) -- (v1) -- (v3) -- (f4),
-          (f2) -- (v1) --[scalar] (v2) -- (v3) -- (f5),
-          (f3) -- (v2) -- (f6), };
-      \end{feynman}
-    \end{tikzpicture}}
-  - \f12 \shrink{
-    \begin{tikzpicture}
-      \begin{feynman}
-        \vertex (v1);
-        \vertex[below right = of v1] (v2);
-        \vertex[above right = of v2] (v3);
-        \vertex[above left = of v1] (f1);
-        \vertex[left = of v1] (f2);
-        \vertex[below left = of v2] (f3);
-        \vertex[above right = of v3] (f4);
-        \vertex[right = of v3] (f5);
-        \vertex[below right = of v2] (f6);
-        \diagram* {
-          (f1) -- (v1) -- (v3) -- (f4),
-          (f2) -- (v1) -- (v2) --[scalar] (v3) -- (f5),
-          (f3) -- (v2) -- (f6), };
-      \end{feynman}
-    \end{tikzpicture}} \\
-  = \shrink{
-    \begin{tikzpicture}
-      \begin{feynman}
-        \vertex (v1);
-        \vertex[below right = of v1] (v2);
-        \vertex[above right = of v2] (v3);
-        \vertex[above left = of v1] (f1);
-        \vertex[left = of v1] (f2);
-        \vertex[below left = of v2] (f3);
-        \vertex[above right = of v3] (f4);
-        \vertex[right = of v3] (f5);
-        \vertex[below right = of v2] (f6);
-        \diagram* {
-          (f1) -- (v1) --[scalar] (v3) -- (f4),
-          (f2) -- (v1) --[scalar] (v2) --[scalar] (v3) -- (f5),
-          (f3) -- (v2) -- (f6), };
-      \end{feynman}
-    \end{tikzpicture}}
-  - \shrink{
-    \begin{tikzpicture}
-      \begin{feynman}
-        \vertex (v1);
-        \vertex[below right = of v1] (v2);
-        \vertex[above right = of v2] (v3);
-        \vertex[above left = of v1] (f1);
-        \vertex[left = of v1] (f2);
-        \vertex[below left = of v2] (f3);
-        \vertex[above right = of v3] (f4);
-        \vertex[right = of v3] (f5);
-        \vertex[below right = of v2] (f6);
-        \diagram* {
-          (f1) -- (v1) -- (v3) -- (f4),
-          (f2) -- (v1) --[scalar] (v2) -- (v3) -- (f5),
-          (f3) -- (v2) -- (f6), };
-      \end{feynman}
-    \end{tikzpicture}}.
-\end{multline}
-The third-order three-body interaction Hamiltonian is therefore
-\begin{align}
-  H_3^{(3)} &= \shrink{
-    \begin{tikzpicture}
-      \begin{feynman}
-        \vertex (v1);
-        \vertex[below right = of v1] (v2);
-        \vertex[above right = of v2] (v3);
-        \vertex[above left = of v1] (f1);
-        \vertex[left = of v1] (f2);
-        \vertex[below left = of v2] (f3);
-        \vertex[above right = of v3] (f4);
-        \vertex[right = of v3] (f5);
-        \vertex[below right = of v2] (f6);
-        \diagram* {
-          (f1) -- (v1) --[scalar] (v3) -- (f4),
-          (f2) -- (v1) --[scalar] (v2) --[scalar] (v3) -- (f5),
-          (f3) -- (v2) -- (f6), };
-      \end{feynman}
-    \end{tikzpicture}}
-  - \shrink{
-    \begin{tikzpicture}
-      \begin{feynman}
-        \vertex (v1);
-        \vertex[below right = of v1] (v2);
-        \vertex[above right = of v2] (v3);
-        \vertex[above left = of v1] (f1);
-        \vertex[left = of v1] (f2);
-        \vertex[below left = of v2] (f3);
-        \vertex[above right = of v3] (f4);
-        \vertex[right = of v3] (f5);
-        \vertex[below right = of v2] (f6);
-        \diagram* {
-          (f1) -- (v1) -- (v3) -- (f4),
-          (f2) -- (v1) --[scalar] (v2) -- (v3) -- (f5),
-          (f3) -- (v2) -- (f6), };
-      \end{feynman}
-    \end{tikzpicture}}
-   - \shrink{
-    \begin{tikzpicture}
-      \begin{feynman}
-        \vertex (v1);
-        \vertex[right = of v1] (v2);
-        \vertex[below right = of v2] (v3);
-        \vertex[above left = of v1] (f1);
-        \vertex[below left = of v1] (f2);
-        \vertex[below left = of v3] (f3);
-        \vertex[above right = of v2] (f4);
-        \vertex[above right = of v3] (f5);
-        \vertex[below right = of v3] (f6);
-        \diagram* {
-          (f1) -- (v1) --[half left, scalar] (v2) -- (f4),
-          (f2) -- (v1) --[half right, scalar] (v2) -- (v3) -- (f5),
-          (f3) -- (v3) -- (f6), };
-      \end{feynman}
-    \end{tikzpicture}}
-  - \shrink{
-    \begin{tikzpicture}
-      \begin{feynman}
-        \vertex (v1);
-        \vertex[right = of v1] (v2);
-        \vertex[below right = of v2] (v3);
-        \vertex[above left = of v1] (f1);
-        \vertex[below left = of v1] (f2);
-        \vertex[below left = of v3] (f3);
-        \vertex[above right = of v2] (f4);
-        \vertex[above right = of v3] (f5);
-        \vertex[below right = of v3] (f6);
-        \diagram* {
-          (f1) -- (v1) --[half left] (v2) -- (f4),
-          (f2) -- (v1) --[half right] (v2) --[scalar] (v3) -- (f5),
-          (f3) -- (v3) -- (f6), };
-      \end{feynman}
-    \end{tikzpicture}} \nonumber \\[1em]
-  &= \p{\alpha_3^{(3)} - \alpha_5^{(3)}} \H_{3,1}^{(3)}
-  - \p{\alpha_{4,3}^{(3)} + \alpha_5^{(3)}} \H_{3,2}^{(3)},
-  \label{eq:H_3_3}
-\end{align}
-where, letting $E_{mn}\equiv E_m+E_n$,
-\begin{align}
-  \alpha_3^{(3)} \equiv \sum_{\substack{\ell+m>0\\\ell+n>0}}
-  \f{K_{\ell m} K^m_n K_{\ell n}}{E_{\ell m} E_{\ell n}},
+where
+\begin{align}
+  \alpha_{3,2}^{(3)}
+  \equiv \sum_{\substack{\ell+m>0\\n>0}}
+  \f{K_{\ell m} K_n}{E_{\ell m} E_n}
+  \p{K^{\ell m}_n - \f{K_{\ell m} K_n}{K}},
   &&
@@
   \equiv K \sum_{m+n>0} \f{K_{mn}^2}{E_{mn}^2},
-  &&
-  \alpha_5^{(3)}
-  \equiv  K \sum_{n>0} \f{K_n^2}{E_n^2},
-  \label{eq:alph_3_3}
-\end{align}
-\begin{align}
-  \H_{3,1}^{(3)} \equiv \sum_{\abs{\set{\mu,\nu,\rho}}=3}
-  G^{rs}_{r's'} G^{\nu s'\rho t}_{\nu's''\rho't'} G^{r's''}_{r''s'''}
-  \c_{\mu r''}^\dag \c_{\nu's'''}^\dag \c_{\rho't'}^\dag
-  \c_{\rho t} \c_{\nu s} \c_{\mu r}
-\end{align}
+  \label{eq:alph_3_3_2}
+\end{align}
+and
 \begin{align}
@@
 \end{align}
-Note that although \eqref{eq:H_3_3_S} is a loop diagram, the
-associated factors $\alpha_3^{(3)}$ and $\alpha_5^{(3)}$ in
-$H_3^{(3)}$ are finite.  At large $n$, atoms become free particles for
-which $n$ essentially indexes discrete momentum states.  These atoms
-thus have an energy which asymptotically scales as
-$E_n\sim n^2\equiv n_\x^2+n_\y^2+n_\z^2$.  Furthermore, the
-oscillatory behavior of atomic wavefunctions with increasing motional
-state indices $\ell,m$ implies that the overlap integral $K_{\ell m}$
-becomes sharply peaked at $\ell\approx m$ as $\ell$ and $m$ get large.
-The asymptotic behavior of $\alpha_3^{(3)}$ at large $\ell,m,n$ is
-therefore
-\begin{align}
-  \alpha_3^{(3)}
-  \sim \int \f{\d^3\ell~\d^3m~\d^3n}{\p{\ell^2+m^2}\p{\ell^2+n^2}}
-  ~\delta\p{\ell-m}\delta\p{\ell-n}
-  \sim \int \f{\d^3\ell}{\ell^4}
-  \sim \int_{\ell_{\t{min}}}^\infty \f{\d\ell}{\ell^2}
-  \sim \f1{\ell_{\t{min}}},
-  \label{eq:alph_3_3_integral}
-\end{align}
-where in the last integral we changed to spherical coordinates, and
-$\ell_{\t{min}}^2$ is the minimum value of $\ell^2$ for which
-\begin{enumerate*}
-\item the energy $E_\ell\sim\ell^2$, and
-\item the integral expression in \eqref{eq:alph_3_3_integral} is a
-  good approximation to the corresponding sum in \eqref{eq:alph_3_3}.
-\end{enumerate*}
-Note that the introduction of $\ell_{\t{min}}$ amounts to neglecting a
-finite number of terms in the sum over $\ell,m,n$ in
-\eqref{eq:alph_3_3}, whose contribution to the value of
-$\alpha_3^{(3)}$ is finite.  Convergence of $\alpha_5^{(3)}$ is
-similarly guaranteed by the fact that the overlap integral $K_n$ does
-not asymptotically grow with increasing $n$, such that
-$\alpha_5^{(3)}$ asymptotically behaves as
-\begin{align}
-  \alpha_5^{(3)} \sim \int \f{\d^3 n}{n^4}
-  \sim \int_{n_{\t{min}}}^\infty \f{\d n}{n^2}
-  \sim \f1{n_{\t{min}}},
-  \label{eq:alph_5_3_integral}
-\end{align}
-where again $n_{\t{min}}$ is defined similarly to $\ell_{\t{min}}$.
+An equal contribution comes from the mirror images of these diagrams,
+such that the net third-order three-body interaction Hamiltonian is
+\begin{align}
+  H_3^{(3)} = \p{\alpha_{3,1}^{(3)} - \alpha_5^{(3)}} \H_{3,1}^{(3)}
+  + \p{2\alpha_{3,2}^{(3)} - \alpha_{4,3}^{(3)} - \alpha_5^{(3)}}
+  \H_{3,2}^{(3)}.
+  \label{eq:H_3_3}
+\end{align}
+Note that the aforementioned divergence and its cancellation are
+buried in $\alpha_{3,2}^{(3)}$.  Formally, this factor is calculated
+by imposing an ultraviolet cutoff $\Lambda$ for the maximum values of
+motional state indices $\ell,m,n$, and then taking the limit
+$\Lambda\to\infty$.  This procedure ensures that $\alpha_{3,2}^{(3)}$
+always remains finite.
 
@@
 the form
-\begin{align}
-  \begin{tikzpicture}
-    \begin{feynman}
-      \vertex[label=0:$~m\mu q'$] (v1);
-      \vertex[below right = of v1, label=0:$~n\mu'q''$] (v2);
-      \vertex[below right = of v2] (v3);
-      \vertex[above left = of v1] (f1) {$\mu q$};
-      \vertex[below left = of v1] (f2) {$\nu r$};
-      \vertex[below left = of v2] (f3) {$\rho s$};
-      \vertex[below left = of v3] (f4) {$\sigma t$};
-      \vertex[above right = of v1] (f5) {$\nu r'$};
-      \vertex[above right = of v2] (f6) {$\rho's'$};
-      \vertex[above right = of v3] (f7) {$\sigma t'$};
-      \vertex[below right = of v3] (f8) {$\mu'q'''$};
-      \diagram* {
-        (f1) -- (v1) -- (f5),
-        (f2) -- (v1) -- (f5),
-        (v1) --[scalar] (v2) --[scalar] (v3) -- (f8),
-        (f3) -- (v2) -- (f6),
-        (f4) -- (v3) -- (f7),
-      };
-    \end{feynman}
-  \end{tikzpicture}
-  \propto K_m K^m_n K_n
-  G^{qr}_{q'r'} G^{\mu q'\rho s}_{\mu'q''\rho's'} G^{q''t}_{q'''t'}
-  \c_{\mu'q'''}^\dag \c_{\nu r'}^\dag \c_{\rho's'}^\dag \c_{\sigma t'}^\dag
-  \c_{\sigma t} \c_{\rho s} \c_{\nu r} \c_{\mu q},
-  \label{eq:H_4_3_C}
-\end{align}
 \begin{align}
@@
   \c_{\mu q''}^\dag \c_{\nu r''}^\dag \c_{\rho s'}^\dag \c_{\sigma t'}^\dag
-  \c_{\sigma t} \c_{\rho s} \c_{\nu r} \c_{\mu q},
+  \c_{\sigma t} \c_{\rho s} \c_{\nu r} \c_{\mu q}
   \label{eq:H_4_3_B}
 \end{align}
-and the mirror image of \eqref{eq:H_4_3_B}.  As we are computing the
-leading-order contribution to effective four-body interactions, there
-are no counter-terms contributions.  In principle, there is now also
-the possibility to make the disconnected diagrams of the form
+and its mirror image, as well as
+\begin{align}
+  \begin{tikzpicture}
+    \begin{feynman}
+      \vertex[label=0:$~m\mu q'$] (v1);
+      \vertex[below right = of v1, label=0:$~n\mu'q''$] (v2);
+      \vertex[below right = of v2] (v3);
+      \vertex[above left = of v1] (f1) {$\mu q$};
+      \vertex[below left = of v1] (f2) {$\nu r$};
+      \vertex[below left = of v2] (f3) {$\rho s$};
+      \vertex[below left = of v3] (f4) {$\sigma t$};
+      \vertex[above right = of v1] (f5) {$\nu r'$};
+      \vertex[above right = of v2] (f6) {$\rho's'$};
+      \vertex[above right = of v3] (f7) {$\sigma t'$};
+      \vertex[below right = of v3] (f8) {$\mu'q'''$};
+      \diagram* {
+        (f1) -- (v1) -- (f5),
+        (f2) -- (v1) -- (f5),
+        (v1) --[scalar] (v2) --[scalar] (v3) -- (f8),
+        (f3) -- (v2) -- (f6),
+        (f4) -- (v3) -- (f7),
+      };
+    \end{feynman}
+  \end{tikzpicture}
+  \propto K_m K^m_n K_n
+  G^{qr}_{q'r'} G^{\mu q'\rho s}_{\mu'q''\rho's'} G^{q''t}_{q'''t'}
+  \c_{\mu'q'''}^\dag \c_{\nu r'}^\dag \c_{\rho's'}^\dag \c_{\sigma t'}^\dag
+  \c_{\sigma t} \c_{\rho s} \c_{\nu r} \c_{\mu q}.
+  \label{eq:H_4_3_C}
+\end{align}
+As we are computing the leading-order contribution to effective
+four-body interactions, there are no counter-terms contributions.  In
+principle, there is now also the possibility to make the disconnected
+diagrams of the form
 \begin{align}
@@
 
-Observing that the diagrams of the form in \eqref{eq:H_4_3_B} are
-equal to their mirror image, the positive contributions to the
-third-order four-body interaction Hamiltonian are
+The contribution to the third-order four-body interaction Hamiltonian
+from diagrams of the form in \eqref{eq:H_4_3_B} is
 \begin{align}
@@
     \end{tikzpicture}}
-  + \shrink{
-    \begin{tikzpicture}
-      \begin{feynman}
-        \vertex (v1);
+  - \f12 \shrink{
+    \begin{tikzpicture}
+      \begin{feynman}
+        \vertex (v1);
+        \vertex[above right = 1.5em of v1] (v2);
         \vertex[below right = 2.5em of v1] (v3);
-        \vertex[below left = 1.7em of v3] (v2);
         \vertex[above left = of v1] (f1);
-        \vertex[left = of v1] (f2);
-        \vertex[left = of v2] (f3);
-        \vertex[below left = of v2] (f4);
-        \vertex[above right = of v1] (f5);
-        \vertex[above right = of v3] (f6);
-        \vertex[below right = of v3] (f7);
-        \vertex[below right = of v2] (f8);
-        \diagram* {
-          (f1) -- (v1) -- (f5),
-          (f2) -- (v1) --[scalar] (v3) -- (f6),
-          (f3) -- (v2) --[scalar] (v3) -- (f7),
-          (f4) -- (v2) -- (f8), };
+        \vertex[below left = of v1] (f2);
+        \vertex[above left = of v2] (f3);
+        \vertex[below left = of v3] (f4);
+        \vertex[above right = of v2] (f5);
+        \vertex[right = of v2] (f6);
+        \vertex[right = of v3] (f7);
+        \vertex[below right = of v3] (f8);
+        \diagram* {
+          (f1) -- (v1) --[scalar] (v2) -- (f6),
+          (f2) -- (v1) -- (v3) -- (f7),
+          (f3) -- (v2) -- (f5),
+          (f4) -- (v3) -- (f8),
+        };
       \end{feynman}
     \end{tikzpicture}}
-  + \shrink{
+  = \p{\alpha_{4,1}^{(3)} - \f12\alpha_5^{(3)}} \H_{4,1},
+\end{align}
+where
+\begin{align}
+  \alpha_{4,1}^{(3)}
+  \equiv \sum_{\substack{m\ge0\\n>0}} \f{K_{mn} K_m K_n}{E_{mn} E_n},
+\end{align}
+and
+\begin{align}
+  \H_{4,1}^{(3)}
+  \equiv \sum_{\abs{\set{\mu,\nu,\rho,\sigma}}=4}
+  G^{qr}_{q'r'} G^{q's}_{q''s'} G^{r't}_{r''t'}
+  \c_{\mu q''}^\dag \c_{\nu r''}^\dag \c_{\rho s'}^\dag \c_{\sigma t'}^\dag
+  \c_{\sigma t} \c_{\rho s} \c_{\nu r} \c_{\mu q}.
+\end{align}
+An equal contribution comes from the mirror images of these diagrams.
+The contribution from diagrams of the form in \eqref{eq:H_4_3_C},
+meanwhile, is
+\begin{align}
+  \shrink{
     \begin{tikzpicture}
@@
     \end{tikzpicture}}
-  = 2 \shrink{
-    \begin{tikzpicture}
-      \begin{feynman}
-        \vertex (v1);
-        \vertex[above right = 1.5em of v1] (v2);
-        \vertex[below right = 2.5em of v1] (v3);
-        \vertex[above left = of v1] (f1);
-        \vertex[below left = of v1] (f2);
-        \vertex[above left = of v2] (f3);
-        \vertex[below left = of v3] (f4);
-        \vertex[above right = of v2] (f5);
-        \vertex[right = of v2] (f6);
-        \vertex[right = of v3] (f7);
-        \vertex[below right = of v3] (f8);
-        \diagram* {
-          (f1) -- (v1) --[scalar] (v2) -- (f6),
-          (f2) -- (v1) --[scalar] (v3) -- (f7),
-          (f3) -- (v2) -- (f5),
-          (f4) -- (v3) -- (f8),
-        };
-      \end{feynman}
-    \end{tikzpicture}}
-  + \shrink{
+  - \f12 \shrink{
     \begin{tikzpicture}
@@
           (f2) -- (v1) -- (f5),
-          (v1) --[scalar] (v2) --[scalar] (v3) -- (f8),
+          (v1) --[scalar] (v2) -- (v3) -- (f8),
           (f3) -- (v2) -- (f6),
@@
       \end{feynman}
-    \end{tikzpicture}},
-\end{align}
-while the negative contributions are
-\begin{multline}
-  \f12 \shrink{
-    \begin{tikzpicture}
-      \begin{feynman}
-        \vertex (v1);
-        \vertex[above right = 1.5em of v1] (v2);
-        \vertex[below right = 2.5em of v1] (v3);
-        \vertex[above left = of v1] (f1);
-        \vertex[below left = of v1] (f2);
-        \vertex[above left = of v2] (f3);
-        \vertex[below left = of v3] (f4);
-        \vertex[above right = of v2] (f5);
-        \vertex[right = of v2] (f6);
-        \vertex[right = of v3] (f7);
-        \vertex[below right = of v3] (f8);
-        \diagram* {
-          (f1) -- (v1) --[scalar] (v2) -- (f6),
-          (f2) -- (v1) -- (v3) -- (f7),
-          (f3) -- (v2) -- (f5),
-          (f4) -- (v3) -- (f8),
-        };
-      \end{feynman}
     \end{tikzpicture}}
-  + \f12 \shrink{
-    \begin{tikzpicture}
-      \begin{feynman}
-        \vertex (v1);
-        \vertex[below right = 2.5em of v1] (v3);
-        \vertex[below left = 1.7em of v3] (v2);
-        \vertex[above left = of v1] (f1);
-        \vertex[left = of v1] (f2);
-        \vertex[left = of v2] (f3);
-        \vertex[below left = of v2] (f4);
-        \vertex[above right = of v1] (f5);
-        \vertex[above right = of v3] (f6);
-        \vertex[below right = of v3] (f7);
-        \vertex[below right = of v2] (f8);
-        \diagram* {
-          (f1) -- (v1) -- (f5),
-          (f2) -- (v1) -- (v3) -- (f6),
-          (f3) -- (v2) --[scalar] (v3) -- (f7),
-          (f4) -- (v2) -- (f8), };
-      \end{feynman}
-    \end{tikzpicture}}
-  + \f12 \shrink{
+  - \f12 \shrink{
     \begin{tikzpicture}
@@
           (f2) -- (v1) -- (f5),
-          (v1) --[scalar] (v2) -- (v3) -- (f8),
-          (f3) -- (v2) -- (f6),
-          (f4) -- (v3) -- (f7),
-        };
-      \end{feynman}
-    \end{tikzpicture}}
-  + \f12 \shrink{
-    \begin{tikzpicture}
-      \begin{feynman}
-        \vertex (v1);
-        \vertex[below right = of v1] (v2);
-        \vertex[below right = of v2] (v3);
-        \vertex[above left = of v1] (f1);
-        \vertex[below left = of v1] (f2);
-        \vertex[below left = of v2] (f3);
-        \vertex[below left = of v3] (f4);
-        \vertex[above right = of v1] (f5);
-        \vertex[above right = of v2] (f6);
-        \vertex[above right = of v3] (f7);
-        \vertex[below right = of v3] (f8);
-        \diagram* {
-          (f1) -- (v1) -- (f5),
-          (f2) -- (v1) -- (f5),
           (v1) -- (v2) --[scalar] (v3) -- (f8),
@@
       \end{feynman}
-    \end{tikzpicture}} \\
-  = \shrink{
-    \begin{tikzpicture}
-      \begin{feynman}
-        \vertex (v1);
-        \vertex[above right = 1.5em of v1] (v2);
-        \vertex[below right = 2.5em of v1] (v3);
-        \vertex[above left = of v1] (f1);
-        \vertex[below left = of v1] (f2);
-        \vertex[above left = of v2] (f3);
-        \vertex[below left = of v3] (f4);
-        \vertex[above right = of v2] (f5);
-        \vertex[right = of v2] (f6);
-        \vertex[right = of v3] (f7);
-        \vertex[below right = of v3] (f8);
-        \diagram* {
-          (f1) -- (v1) --[scalar] (v2) -- (f6),
-          (f2) -- (v1) -- (v3) -- (f7),
-          (f3) -- (v2) -- (f5),
-          (f4) -- (v3) -- (f8),
-        };
-      \end{feynman}
     \end{tikzpicture}}
-  + \shrink{
-    \begin{tikzpicture}
-      \begin{feynman}
-        \vertex (v1);
-        \vertex[below right = of v1] (v2);
-        \vertex[below right = of v2] (v3);
-        \vertex[above left = of v1] (f1);
-        \vertex[below left = of v1] (f2);
-        \vertex[below left = of v2] (f3);
-        \vertex[below left = of v3] (f4);
-        \vertex[above right = of v1] (f5);
-        \vertex[above right = of v2] (f6);
-        \vertex[above right = of v3] (f7);
-        \vertex[below right = of v3] (f8);
-        \diagram* {
-          (f1) -- (v1) -- (f5),
-          (f2) -- (v1) -- (f5),
-          (v1) --[scalar] (v2) -- (v3) -- (f8),
-          (f3) -- (v2) -- (f6),
-          (f4) -- (v3) -- (f7),
-        };
-      \end{feynman}
-    \end{tikzpicture}},
-\end{multline}
-such that the net third-order four-body interaction Hamiltonian is
-\begin{align}
-  H_4^{(3)} &= 2 \shrink{
-    \begin{tikzpicture}
-      \begin{feynman}
-        \vertex (v1);
-        \vertex[above right = 1.5em of v1] (v2);
-        \vertex[below right = 2.5em of v1] (v3);
-        \vertex[above left = of v1] (f1);
-        \vertex[below left = of v1] (f2);
-        \vertex[above left = of v2] (f3);
-        \vertex[below left = of v3] (f4);
-        \vertex[above right = of v2] (f5);
-        \vertex[right = of v2] (f6);
-        \vertex[right = of v3] (f7);
-        \vertex[below right = of v3] (f8);
-        \diagram* {
-          (f1) -- (v1) --[scalar] (v2) -- (f6),
-          (f2) -- (v1) --[scalar] (v3) -- (f7),
-          (f3) -- (v2) -- (f5),
-          (f4) -- (v3) -- (f8),
-        };
-      \end{feynman}
-    \end{tikzpicture}}
-  - \shrink{
-    \begin{tikzpicture}
-      \begin{feynman}
-        \vertex (v1);
-        \vertex[above right = 1.5em of v1] (v2);
-        \vertex[below right = 2.5em of v1] (v3);
-        \vertex[above left = of v1] (f1);
-        \vertex[below left = of v1] (f2);
-        \vertex[above left = of v2] (f3);
-        \vertex[below left = of v3] (f4);
-        \vertex[above right = of v2] (f5);
-        \vertex[right = of v2] (f6);
-        \vertex[right = of v3] (f7);
-        \vertex[below right = of v3] (f8);
-        \diagram* {
-          (f1) -- (v1) --[scalar] (v2) -- (f6),
-          (f2) -- (v1) -- (v3) -- (f7),
-          (f3) -- (v2) -- (f5),
-          (f4) -- (v3) -- (f8),
-        };
-      \end{feynman}
-    \end{tikzpicture}}
-  + \shrink{
-    \begin{tikzpicture}
-      \begin{feynman}
-        \vertex (v1);
-        \vertex[below right = of v1] (v2);
-        \vertex[below right = of v2] (v3);
-        \vertex[above left = of v1] (f1);
-        \vertex[below left = of v1] (f2);
-        \vertex[below left = of v2] (f3);
-        \vertex[below left = of v3] (f4);
-        \vertex[above right = of v1] (f5);
-        \vertex[above right = of v2] (f6);
-        \vertex[above right = of v3] (f7);
-        \vertex[below right = of v3] (f8);
-        \diagram* {
-          (f1) -- (v1) -- (f5),
-          (f2) -- (v1) -- (f5),
-          (v1) --[scalar] (v2) --[scalar] (v3) -- (f8),
-          (f3) -- (v2) -- (f6),
-          (f4) -- (v3) -- (f7),
-        };
-      \end{feynman}
-    \end{tikzpicture}}
-  - \shrink{
-    \begin{tikzpicture}
-      \begin{feynman}
-        \vertex (v1);
-        \vertex[below right = of v1] (v2);
-        \vertex[below right = of v2] (v3);
-        \vertex[above left = of v1] (f1);
-        \vertex[below left = of v1] (f2);
-        \vertex[below left = of v2] (f3);
-        \vertex[below left = of v3] (f4);
-        \vertex[above right = of v1] (f5);
-        \vertex[above right = of v2] (f6);
-        \vertex[above right = of v3] (f7);
-        \vertex[below right = of v3] (f8);
-        \diagram* {
-          (f1) -- (v1) -- (f5),
-          (f2) -- (v1) -- (f5),
-          (v1) --[scalar] (v2) -- (v3) -- (f8),
-          (f3) -- (v2) -- (f6),
-          (f4) -- (v3) -- (f7),
-        };
-      \end{feynman}
-    \end{tikzpicture}} \nonumber \\
-  &= \p{2\alpha_{4,1}^{(3)} - \alpha_5^{(3)}} \H_{4,1}^{(3)}
-  + \p{\alpha_{4,2}^{(3)} - \alpha_5^{(3)}} \H_{4,2}^{(3)},
-  \label{eq:H_4_3}
+  = \p{\alpha_{4,2}^{(3)} - \alpha_5^{(3)}} \H_{4,2},
 \end{align}
@@
 \begin{align}
-  \alpha_{4,1}^{(3)}
-  \equiv \sum_{\substack{m\ge0\\n>0}} \f{K_{mn} K_m K_n}{E_{mn} E_n},
-  &&
   \alpha_{4,2}^{(3)}
   \equiv \sum_{m,n>0} \f{K_m K^m_n K_n}{E_m E_n},
-  &&
-  \alpha_5^{(3)}
-  \equiv  K \sum_{n>0} \f{K_n^2}{E_n^2},
-\end{align}
-\begin{align}
-  \H_{4,1}^{(3)}
-  \equiv \sum_{\abs{\set{\mu,\nu,\rho,\sigma}}=4}
-  G^{qr}_{q'r'} G^{q's}_{q''s'} G^{r't}_{r''t'}
-  \c_{\mu q''}^\dag \c_{\nu r''}^\dag \c_{\rho s'}^\dag \c_{\sigma t'}^\dag
-  \c_{\sigma t} \c_{\rho s} \c_{\nu r} \c_{\mu q},
-\end{align}
+\end{align}
+and
 \begin{align}
@@
   \c_{\sigma t} \c_{\rho s} \c_{\nu r} \c_{\mu q}.
+\end{align}
+The net third-order four-body interaction Hamiltonian is therefore
+\begin{align}
+  H_4^{(3)}
+  = \p{2\alpha_{4,1}^{(3)} - \alpha_5^{(3)}} \H_{4,1}^{(3)}
+  + \p{\alpha_{4,2}^{(3)} - \alpha_5^{(3)}} \H_{4,2}^{(3)}.
+  \label{eq:H_4_3}
 \end{align}
@@
   U_{3,\g}^{(3)}
-  = \p{\alpha_3^{(3)} - \alpha_5^{(3)}} 2 G_\g^3
-  - \p{\alpha_{4,3}^{(3)} + \alpha_5^{(3)}} G_\g^3,
+  = \p{\alpha_{3,1}^{(3)} - \alpha_5^{(3)}} 2 G_\g^3
+  + \p{2\alpha_{3,2}^{(3)} - \alpha_{4,3}^{(3)} - \alpha_5^{(3)}} G_\g^3,
 \end{align}
@@
   U_{3,+}^{(3)}
-  &= \p{\alpha_3^{(3)} - \alpha_5^{(3)}} \p{G_+^3 + 4 G_+^2 G_\g
+  &= \p{\alpha_{3,1}^{(3)} - \alpha_5^{(3)}} \p{G_+^3 + 4 G_+^2 G_\g
     + G_+ G_-^2 + G_+ G_\g^2 + G_-^3 + G_-^2 G_\g} \nonumber \\
-  &\quad - \p{\alpha_{4,3}^{(3)} + \alpha_5^{(3)}}
+  &\quad + \p{2\alpha_{3,2}^{(3)} - \alpha_{4,3}^{(3)} - \alpha_5^{(3)}}
   \p{G_+^3 + G_+^2 G_\g + G_+ G_-^2 + G_+ G_\g^2 + G_-^2 G_\g},
@@
   U_{3,-}^{(3)}
-  &= \p{\alpha_3^{(3)} - \alpha_5^{(3)}} G_- \p{3 G_+^2
+  &= \p{\alpha_{3,1}^{(3)} - \alpha_5^{(3)}} G_- \p{3 G_+^2
     + 2 G_+ G_- + 8 G_+ G_\g + 3 G_- G_\g + G_\g^2} \nonumber \\
-  &\quad - \p{\alpha_{4,3}^{(3)} + \alpha_5^{(3)}}
+  &\quad + \p{2\alpha_{3,2}^{(3)} - \alpha_{4,3}^{(3)} - \alpha_5^{(3)}}
   G_- \p{3 G_+^2 + 2 G_+ G_- + 2 G_+ G_\g + G_-^2 + G_\g^2},
@@
   &= \p{2\alpha_{4,1}^{(3)} - \alpha_5^{(3)}}
-  2 G_+ G_\g \p{G_+ + G_\g} \nonumber \\
-  &\quad + \p{\alpha_{4,2}^{(3)} - \alpha_5^{(3)}}
+  2 G_+ G_\g \p{G_+ + G_\g}
+  + \p{\alpha_{4,2}^{(3)} - \alpha_5^{(3)}}
   G_+ \p{G_+^2 + 2 G_+ G_\g + 5 G_\g^2},
@@
   &&
-  \alpha_3^{(3)} \equiv \sum_{\substack{\ell+m>0\\\ell+n>0}}
+  \alpha_5^{(3)}
+  \equiv  K \sum_{n>0} \f{K_n^2}{E_n^2},
+\end{align}
+\begin{align}
+  \alpha_{3,1}^{(3)} \equiv \sum_{\substack{\ell+m>0\\\ell+n>0}}
   \f{K_{\ell m} K^m_n K_{\ell n}}{E_{\ell m} E_{\ell n}},
+  &&
+  \alpha_{3,2}^{(3)}
+  \equiv \sum_{\substack{\ell+m>0\\n>0}}
+  \f{K_{\ell m} K_n}{E_{\ell m} E_n}
+  \p{K^{\ell m}_n - \f{K_{\ell m} K_n}{K}},
 \end{align}
@@
   \alpha_{4,3}^{(3)}
-  \equiv K \sum_{m+n>0} \f{K_{mn}^2}{E_{mn}^2},
-\end{align}
-\begin{align}
-  \alpha_5^{(3)}
-  \equiv  K \sum_{n>0} \f{K_n^2}{E_n^2}.
+  \equiv K \sum_{m+n>0} \f{K_{mn}^2}{E_{mn}^2}.
 \end{align}

### 2019-02-27 multi_body_interactions/scripts/interaction_num_methods.py

@@ -45,7 +45,7 @@ def energy_correction_coefficients(lattice_depths, site_number,
 
     # compute all 1-D band energies, spatial wavefunction overlaps, and tunneling rates
     band_energies = np.zeros((3,bands))
-    K_1D = np.zeros((3,bands,bands))
+    K_1D = np.zeros((3,bands,bands,bands))
     if pt_order > 2:
         K_t_1D = np.zeros((3,bands,bands))
         t_1D = np.zeros((3,bands))
@@ -53,7 +53,7 @@ def energy_correction_coefficients(lattice_depths, site_number,
         # if any of the lattice depths are the same, we can recycle our calculations
         if axis > 0 and lattice_depths[axis] == lattice_depths[axis-1]:
             band_energies[axis,:] = band_energies[axis-1,:]
-            K_1D[axis,:,:] = K_1D[axis-1,:,:]
+            K_1D[axis,:,:,:] = K_1D[axis-1,:,:,:]
             if pt_order > 2:
                 K_t_1D[axis,:,:] = K_t_1D[axis-1,:,:]
                 t_1D[axis,:] = t_1D[axis-1,:]
@@ -70,14 +70,16 @@ def energy_correction_coefficients(lattice_depths, site_number,
                 t_1D[axis,nn] = tunneling_1D(lattice_depths[axis],
                                              momenta, fourier_vecs, nn)
             for mm in range(nn,bands):
-                K_1D[axis,nn,mm] = pair_overlap_1D(momenta, fourier_vecs, nn, mm)
-                K_1D[axis,mm,nn] = K_1D[axis,nn,mm]
                 if pt_order > 2:
                     K_t_1D[axis,nn,mm] = pair_overlap_1D(momenta, fourier_vecs, nn, mm,
                                                          neighbors = 1)
                     K_t_1D[axis,mm,nn] = K_t_1D[axis,nn,mm]
 
-    K = np.prod(K_1D[:,0,0])
+                for ll in range(bands):
+                    K_1D[axis,nn,mm,ll] = pair_overlap_1D(momenta, fourier_vecs, nn, mm, ll)
[…]

### 2019-02-27 multi_body_interactions/papers/theory/reviews_NJP/final_reviews/referee_response_2.tex
first draft of second round of review for multi-body paper in NJP

--- 2019-02-14 multi_body_interactions/papers/theory/reviews_NJP/referee_response_2.tex
@@
 
-We thank the referee for taking the time to carefully read our
-manuscript, and for providing thoughtful and constructive feedback.
-Below, we address the comments and concerns brought up by the referee.
-We hope that the referee finds our responses and revisions
-satisfactory, deeming our manuscript acceptable for publication in New
-Journal of Physics.
+We thank the referee for taking the time to read and reconsider our
+revised manuscript.  Concerning the comments:
 
-Excerpts from the referee reports are written in \blue{blue}, excerpts
-from our original manuscript are written in \red{red}, and excerpts
-from the revised version of our manuscript is written in
-\green{green}.  All page, line, equation, and reference numbers
-generally refer to those in our original manuscript, unless explicitly
-stated otherwise.
+\blue{I read the replies from the authors and other report and
+  recommend the publication.}
 
-\begin{enumerate}
-\item Concerning:
+\blue{There are a few discussions on ``renormalization'' schemes in
+  the article and in Referee 1's report.  Let me comment on the scheme
+  in the manuscript. The scheme used here (more precisely
+  ``regularization'' scheme) is a quite standard and has been commonly
+  used in many-body studies since the beginning of many-body
+  diagrammatic calculations. It is based on a k-space ``pseudo''
+  potential.  From the point of view of low energy few-body physics
+  studied here, it is indeed equivalent to the other well known
+  standard real space ``differential pseudo'' potential Referee A
+  might have in mind (I speculate that what he/she meant by ``pseudo
+  potential''). But more importantly, it can be easily and
+  straightforwardly generalized to arbitrary dimensions and can be
+  integrated into studies of complex/sophisticated collective
+  many-body effects associated with Fermi surfaces and/or sound waves
+  in symmetry breaking states.}
 
-  \blue{Largely motivated by the experimental spectroscopy data on
-    87Sr atoms, the authors of manuscript NJP-109533 have studied
-    effective few-body interactions in mesoscopic Sr atomic clusters
-    confined by optical lattice potentials and applied the effective
-    Hamiltonian to characterize the spectra of interacting Sr atoms.
-    The authors have also compared the result of theory with data and
-    shown general consistency.}
-
-  \blue{The main results are summarized in section III and IV.  In
-    section II, the authors presented the detailed discussions on the
-    effective interactions via the standard Schrieffer-Wolff
-    transformation. The low energy manifold where the orbital motion
-    is fixed in the ground state is separated from excited manifolds
-    involving excited orbitals. The low energy physics therefore is
-    purely driven by the internal SU(N) spin dynamics and the excited
-    orbitals will further induce few-body interaction between SU(N)
-    spins.  These can be systematically studied at least in the weakly
-    interacting systems or in the perturbative limit where the single
-    particle orbital motion is well gapped and Schrieffer-Wolff
-    transformation is applicable. In section IV, the authors benched
-    their numerical results against the data on 87Sr and emphasized
-    the situation where the induced effective interactions might play
-    a substantial role.}
-
-  \blue{I find the article is well written and results have been
-    carefully organized and are well presented.  Although at
-    qualitatively level, these higher order effects are relatively
-    small (not surprising as perturbative effects) and do not modify
-    the results in a drastic way, they do offer better agreements with
-    experimental data. And in the extreme cases in Fig.5 (3+, 4+, 5+
-    lines), numerical values of these effects o(G\^{}3) can be as
-    larger as 10-20\% (to my bare eyes) although in this limit it
-    becomes less clear whether the contributions not included are also
-    substantial.}
-
-  \blue{A few suggestions the authors can take into account to improve
-    the presentation.}
-
-  We thank the referee for their summary of our work, and for
-  remarking that our ``article is well written'', and that our
-  ``results have been carefully organized are well presented''.  We
-  hope that the revisions we outline below address the referee's
-  suggestions to a satisfactory degree.
-
-\item Concerning:
-
-  \blue{1) In the caption of Fig 1, the authors shall include the
-    dashed solid line definitions to make it self-contained.}
-
-  We thank the referee for the suggestion to make figure 1 more
-  self-contained.  Following this suggestion, we have added the
-  definition of solid and dashed lines to the caption of figure 1, as
-  well as a sentence describing how to translate a diagram into a
-  sequence of operators.  In total, we have replaced the text (page
-  11, line 13):
-
-  \red{... $n>0$ and $\c_{\mu s}\equiv\c_{0,\mu s}$.  For the sake
-    presentation, ...}
-
-  by:
-
-  \green{... $n>0$ and $\c_{\mu s}\equiv\c_{0,\mu s}$.  Diagrams are
-    read from left to right to construct a sequence of operators from
-    right to left.  Solid (dashed) lines represent field operators
-    acting on the lowest (arbitrary) motional states.  For the sake
-    presentation, ...}
-
-\item Concerning the closely related comments/questions::
-
-  \blue{2) Section IV, the author shall add discussions about why some
-    states are more affected than the others by the 4-body interaction
-    etc.  Data are clearly presented and well compared with
-    experiments but there seems to be a lack of central message. What
-    is the main lesson? After all, they are high order effects so
-    naively and logically one would say they play little role in the
-    system. But on other hand, numerically in some cases they appear
-    to be amplified up to 20\%, not so small anymore.  Is it
-    accidental for 3+, 4+, 5+ states or there are some more
-    fundamental reasons for that?}
-
-  \blue{3) Related to 2), when the correction becomes 10-20\%, shall
-    we expect that higher order effects of order of o(G\^{}4) etc also
-    become important?}
-
-  \label{pt:figs}
-
-  We thank the referee for their careful reading of our results in
-  section IV, and for pointing out that our manuscript would greatly
-  benefit from a deeper discussion of figures 4 and 5.  The referee
-  asks questions that follow naturally from an inspection of these
-  figures; our manuscript should address these questions directly.
-
-  Indeed, there is a good reason why multi-body interaction energy
-  corrections are more prominent for the symmetric ($3+$, $4+$, $5+$)
-  states than for the asymmetric ($3-$, $4-$, $5-$) ones.
-  Specifically, there is a competition between contributions of
-  opposite sign to the interaction energy of asymmetric states.
-  Concerning higher-order effects, it is not obvious when exactly
-  further corrections ($\sim G^4$) should become relevant.  Computing
-  higher-order corrections requires more systematic methods than those
-  used in our manuscript, and we are simply unsure whether considering
-  these corrections is worthwhile without first performing a more
-  detailed analysis of systematic errors in both theory and
-  experiment.
-
-  To address these questions and add a more pointed discussion of the
-  lessons from figures 4 and 5, we have rearranged, rewritten, and
-  added new text to the following two paragraphs about these figures
-  (page 23, line 54):
-
-  \red{Identifying peaks in excitation spectra such as in figure 3
-    constitutes a measurement of many-body excitation energies, which
-    was performed in ref.~[34] to detect signatures of effective
-    multi-body interactions.  Figure 4 shows a comparison between
-    experimental measurements of the many-body excitation energies
-    $\Delta_{NX}$ for $N\in\set{3,4,5}$ at various lattice depths
-    $\U$, and the corresponding values of $\Delta_{NX}$ predicted by
-    the low-energy effective theory at different orders in the
-    coupling constants.  A summary of the data in figure 4 is provided
-    in figure 5.  These figures show a clear improvement of the
-    low-energy effective theory we have developed with increasing
-    order in the coupling constants, in particular demonstrating the
-    need to consider multi-body interactions in a single-band model of
-    the atoms without motional excitations.}
-
-  \red{We expect that the dominant source of error in the third-order
-    effective theory comes from the fact that we neglect the
-    inter-site matrix elements of all Hamiltonians.  This source of
-    error is discussed in Appendix E, and leads to theoretical
-    uncertainties represented by error bars in figure 4.  As these
-    uncertainties are sometimes unable to account for the residual
-    disagreement between experimental measurements and the effective
-    theory, better agreement requires either a better understanding of
-    experimental procedures and error, a better understanding of
-    theoretical uncertainties, or consideration of yet higher-order
-    terms in the effective theory.  Due to a combinatorial explosion
-    in the number of diagrams which appear at increasing perturbative
-    orders of the effective theory, we need more systematic methods to
-    compute effective multi-body Hamiltonians at fourth order; we
-    therefore leave this calculation for future work.}
-
-  which now read:
-
-  \green{Identifying peaks in excitation spectra such as in figure 3
-    constitutes a measurement of many-body excitation energies, which
-    was performed in ref.~[34] to detect signatures of effective
-    multi-body interactions.  Figure 4 shows a comparison between (i)
-    experimental measurements of the many-body excitation energies
-    $\Delta_{NX}$ for all $\p{N,X}\in\set{3,4,5}\times\set{+,-}$ at
-    various mean lattice depths $\U$, and (ii) the corresponding
-    values of $\Delta_{NX}$ predicted by the low-energy effective
-    theory at different orders in the coupling constants.  A known
-    source of error in our effective theory comes from neglecting the
-    inter-site matrix elements of all Hamiltonians.  This error is
-    discussed in Appendix E, and leads to theoretical uncertainties
-    represented by error bars on the $\O\p{G^3}$ theory in figure 4.
-    A summary of figure 4 is provided in figure 5.  We note that
-    many-body interaction energy shifts are smaller for asymmetric
-    ($-$) states than symmetric ($+$) ones due to the competition
-    between contributions of opposite sign in the asymmetric case (see
-    rows 2 and 3 of table I).}
-
-  \green{The results in figures 4 and 5 highlight a few important
-    points about ultracold, high-density ${}^{87}$Sr experiments and
-    our low-energy effective theory.  First, these experiments exhibit
-    clear signatures of multi-body interactions, as evidenced by a
-    stark disagreement between the observed many-body excitation
-    energies $\Delta_{NX}$ and those that are predicted by the
-    two-body $\O\p{G}$ theory.  Multi-body interactions are thus
-    crucial for understanding these high-density experiments in the
-    context of a single-band Hubbard model, which naturally arises in
-    the zero-temperature limit when all atoms occupy their motional
-    ground state.  Second, the inter-atomic interactions in these
-    experiments are strong enough to require going beyond leading
-    order for the description of multi-body interactions in the
-    low-energy effective theory.  The formally small quantities
-    organizing our perturbation theory are two-body ground-state
-    interaction energies (proportional to the couplings $G_X$) divided
-    by the spectral gap of the single-atom Hamiltonian $H_0$.  These
-    reduced (dimensionless) interaction energies vary from
-    $\sim0.05-0.15$ in the parameter regimes of the ${}^{87}$Sr
-    experiments considered here (see Appendix F).  As experiments
-    begin to work with more atoms and different lattice depths,
-    therefore, reliably predicting interaction energies may require
-    going to yet higher orders in perturbation theory.  Due to a
-    combinatorial explosion of the number of diagrams which appear at
-    increasing orders in the effective theory, however, we need more
-    systematic methods to compute effective multi-body Hamiltonians at
-    fourth order.  In any case, we are agnostic as to whether such a
-    calculation would provide better agreement between experiment and
-    theory without first performing a detailed analysis of systematic
-    errors.}
-
-  Note that the Appendix F referenced above is not the same as that in
-  the original version of our manuscript, but a new appendix discussed
-  in point \ref{pt:pert_params} below.  We hope that the new text
-  above adequately addresses the questions brought up by the referee,
-  and provides a clear central message for the results in figures 4
-  and 5.
-
-\item Concerning:
-
-  \blue{4) At a phenomenological level, the author shall describe, in
-    connection to the experiment, the range of the perturbation
-    parameter [ a\_X K/Delta] that measures the two-particle
-    interaction energy. For Sr atoms with typical background
-    scattering lengths, this is usually small but does depend on the
-    lattice depth quite sensitively.  The authors shall provide the
-    typical numerical range of this dimensionless interaction energy
-    for the optical lattices (up to 70E\_R in Fig 5) they studied (and
-    experimentally studied). This will help readers to better
-    appreciate their numerical results (surprising or anticipated).}
-
-  \label{pt:pert_params}
-
-  We thank the referee for pointing out that we had overlooked this
-  important bit of information for interpreting the numerical results
-  in our manuscript.  Consequently, we now provide typical values for
-  our perturbative parameters in the discussion of figures 4 and 5
-  (see point \ref{pt:figs} above), and we have added the following new
-  Appendix to our manuscript:
-
-  \setcounter{figure}{0}
-  \renewcommand\thefigure{F.\arabic{figure}}
-
-  \begin{figure}
-    \centering
-    \includegraphics{../figures/perturbative_params.pdf}
-    \caption{\footnotesize Dependence of the perturbative parameters
-      $KG_X/\Delta$ on the lattice depth $\U$.}
-  \end{figure}
-
-  \green{{\bf Appendix F: Perturbative parameters for the effective
-      theory}}
-
-  \green{The perturbative effective theory developed in Section III is
-    organized in powers of the coupling constants $G_X$.  The formally
-    small, dimensionless quantities for this perturbation theory are
-    the two-body interaction energies $K G_X$, where $K$ is a
-    ground-state two-body overlap integral and $G_X$ is a coupling
-    constant, divided by the spectral gap $\Delta$ of the
-    non-interacting Hamiltonian $H_0$.  The quantities $K$, $G_X$, and
-    $\Delta$ all depend on the lattice depth $\U$.  Figure F.1 shows
-    these parameters for the case of ${}^{87}$Sr atoms with
-    $X\in\set{\g\g, \e\g_-, \e\g_+, \e\e}$ at lattice depths
-    $\U\in\sp{30,80}E_{\t{R}}$, where
-    $E_{\t{R}}\approx3.5\times2\pi~\t{kHz}$ is the lattice photon
-    recoil energy of the atoms.  The fact that these perturbative
-    parameters grow with increasing lattice depth $\U$ is a
-    consequence of the fact that the overlap integral $K$ grows faster
-    with $\U$ than the spectral gap $\Delta$.  In the case of a
-    harmonic trap with angular frequency $\omega$, for example, these
-    parameters would be
-    \begin{multline*}
-      \f{K_{\t{HO}} G_X}{\omega}
-      = \f{G_X}{\omega} \int \d^3 x~\abs{\phi_{\t{HO}}}^4
-      = \f{G_X}{\omega} \sp{\int \d x~
-        \abs{\p{\f{\mA\omega}{\pi}}^{1/4} e^{-\mA\omega x^2/2}}^4}^3 \\
-      \sim \f1{\omega}
-      \sp{\omega \int\f{\d\p{\sqrt\omega x}}{\sqrt\omega}
-        ~\abs{e^{-\mA\omega x^2/2}}^4}^3
-      \propto \sqrt{\omega},
-      \tag{F1}
-    \end{multline*}
-    where we assumed that the coupling constants $G_X$ vary weakly
-    with $\omega$.  While this result may seem to suggest that the
-    low-energy effective theory should become better at smaller
-    lattice depths, smaller lattice depths also result in increased
-    theoretical uncertainties from the growing relevance of the
-    inter-site effects discussed in Appendix E.}
-
-  We hope that these additions help improve readers' interpretations
-  of the numerical results in our manuscript.
-
-\item Finally:
-
-  \blue{Overall, the manuscript contains substantial and solid results
-    and data on few-body spectra of SU(N) atoms and can be quite
-    useful to future studies on these atoms. I recommend the
-    publication of the manuscript in NJP after my concerns above are
-    addressed.}
-
-  We thank the referee for remarking that our manuscript ``contains
-  substantial and results'', and hope that our results will indeed be
-  ``quite useful to future studies''.  We hope that the revisions we
-  outlined above have addressed the referee's concerns and suggestions
-  to a satisfactory degree,
-
-\end{enumerate}
+We thank the referee for their recommendation to publish our
+manuscript, and for their input to help clarify a point of confusion
+in our exchanges with referee 1.  While our original response to
+referee 1 was imprecise in conflating the term ``renormalization
+scheme'' with ``regularization scheme'', we have gone through to make
+sure that our manuscript uses these terms properly.
 

### 2019-03-03 multi_body_interactions/papers/theory/reviews_NJP/final_reviews/referee_response_2.tex
final changes to multi-body paper for 2nd round of review

@@
 We thank the referee for their recommendation to publish our
-manuscript, and for their input to help clarify a point of confusion
-in our exchanges with referee 1.  While our original response to
-referee 1 was imprecise in conflating the term ``renormalization
-scheme'' with ``regularization scheme'', we have gone through to make
-sure that our manuscript uses these terms properly.
+manuscript, and for their input on the regularization scheme in our
+manuscript.  In particular, we thank the referee for pointing out that
+our approach to regularizing the two-body delta-function interaction
+is a standard technique in many-body studies, equivalent to another
+common technique based on a regularized real-space ``differential
+pseudo-potential''.
 

### 2019-03-03 multi_body_interactions/papers/theory/paper.tex
some changes to latest renormalization derivation
final changes to multi-body paper for 2nd round of review

@@
 
-At second order in the coupling constants, the expansion for
-$H_{\t{int}}^{(2)}$ in \eqref{eq:H_int_1_2} gives rise to effective
-three-body interactions.  The virtual state of three-body terms in
-$H_{\t{int}}^{(2)}$ cannot have two atoms in excited motional states,
-as otherwise the second application of $H_{\t{int}}$ in
-$H_{\t{int}}^{(2)}$ would have to address both of theses atoms to
-bring them back down to the ground state, resulting in a two-body
-process as in the second diagram of \eqref{eq:H_2_expansion}.  All
-second-order three-body terms must therefore have only one excited
-atom in the virtual state, and take the form
+Our theory of effective multi-body interactions assumes no
+non-universal contribution to the three-body interaction energy, which
+is to say that we assume the absence of real (as opposed to
+effective), bare three-body interactions.  Consequently, three-body
+interactions do not appear until second order in the coupling
+constants of the effective theory, in the expansion of
+$H_{\t{int}}^{(2)}$ in \eqref{eq:H_int_1_2}.  The virtual state of
+three-body terms in $H_{\t{int}}^{(2)}$ cannot have two atoms in
+excited motional states, as otherwise the second application of
+$H_{\t{int}}$ in $H_{\t{int}}^{(2)}$ would have to address both of
+theses atoms to bring them back down to the ground state, resulting in
+a two-body process as in the second diagram of
+\eqref{eq:H_2_expansion}.  All second-order three-body terms must
+therefore have only one excited atom in the virtual state, and take
+the form
 \begin{align}
@@
 \end{align}
-which can be expanded and solved explicitly for the counter-terms to
-find
-\begin{align}
-  \widetilde G^{rs}_{r''s''}
-  = \sum_{\ell,m,r',s'} \f{K_{\ell m}^2}{K E_{\ell m}}
+which can be expanded to find
+\begin{align}
+  K \widetilde G^{rs}_{r''s''}
+  = \sum_{\ell,m,r',s'} \f{K_{\ell m}^2}{E_{\ell m}}
   G^{rs}_{r's'} G^{r's'}_{r''s''}.
@@
 motional state indices $\ell,m,n$, and then taking the limit
-$\Lambda\to\infty$.  This procedure ensures that $\alpha_{3,2}^{(3)}$
-always remains finite.
+$\Lambda\to\infty$.  This procedure ensures that there are no
+divergences in $\alpha_{3,2}^{(3)}$.
 

### 2019-03-18 soc_squeezing/collective_paper/collective_spin.tex
updates (largely wording/symbol changes) to the collective spin paper

--- 2018-12-11 soc_squeezing/paper/collective_spin.tex
@@
 \usepackage{graphicx} % for figures
+\usepackage[caption=false]{subfig} % subfigures (via \subfloat[]{})
+\newcommand{\sref}[1]{\protect\subref{#1}} % for referencing subfigures
 \graphicspath{{./figures/}} % set path for all figures
@@
 
-\title{Short-time simulation of open collective quantum spin systems}
+\title{Structure constants and short-time simulations of open
+  collective quantum spin systems}
 
@@
 sensors\cite{jones2009magnetic}.  These implementations have triggered
-intensive investigations of rich phenomena, including studies of
-quantum criticality\cite{latorre2005entanglement,
-  alcalde2007functional, morrison2008dissipationdriven,
-  morrison2008collective, sarandy2009classical, wang2012quantum,
-  kessler2012dissipative, majd2014lmg}, non-equilibrium
+studies of a variety of rich subjects, including quantum
+criticality\cite{latorre2005entanglement, alcalde2007functional,
+  morrison2008dissipationdriven, morrison2008collective,
+  sarandy2009classical, wang2012quantum, kessler2012dissipative,
+  majd2014lmg}, non-equilibrium
 phenomena\cite{walls1978nonequilibrium, morrison2008dynamical,
@@
 states typically requires nonlinear dynamics that are realized using
-long-range collisional, photon-, or phonon-mediated interactions.
+collisional, photon-mediated, or phonon-mediated interactions.
 Although a truly collective spin model requires uniform, all-to-all
@@
 long time to converge, as simulations become dominated by incoherent
-jumps that generate large numbers of distinct trajectories that need
-to be averaged.
+jumps that generate large numbers of distinct quantum trajectories
+that need to be averaged to compute expectation values.
 
@@
 the equations of motion for Heisenberg operators.  Evaluating this
-expansion requires use of the structure factors of a collective spin
-operator algebra; the calculation of these structure factors is one of
-the main technical results of this work, which we hope will enable
-future studies to develop new analytical insights into the structure
-and dynamics of collective spin systems.
+expansion requires using the structure constants of a collective spin
+operator algebra; the calculation of these structure constants is one
+of the main technical results of this work, which we hope will enable
+future studies to develop new analytical insights into the dynamics of
+collective spin systems.
 
@@
 $s_\pm\equiv s_\x\pm is_\y=\sigma_\pm$ with Pauli operators
-$\sigma_n$, we denote an operator which acts with $s_n$ on the spin
+$\sigma_n$, we denote an operator that acts with $s_n$ on the spin
 indexed by $j$ and trivially (i.e.~with the identity $\1$) on all
@@
 $\gamma_{\J_-}=\gamma_-$.  The commutator in Eq.~\eqref{eq:EOM} can be
-determined from the structure factors of $f_{\v n\v m\v\ell}$ of the
+determined from the structure constants $f_{\v n\v m\v\ell}$ of the
 general product
-$\S_{\v n}\S_{\v m}=\sum_{\v\ell}f_{\v n\v m\v\ell}S_{\v\ell}$ worked
+$\S_{\v n}\S_{\v m}=\sum_{\v\ell}f_{\v n\v m\v\ell}\S_{\v\ell}$ worked
 out in Appendices \ref{sec:identities}--\ref{sec:prod_general}, while
@@
 number of expectation values $\bk{\S_{\v m}}_{t=0}$ and matrix
-elements $T_{\v m\v n}$ that we need to explicitly compute roughly in
-half.
+elements $T_{\v m\v n}$ that we may need to explicitly compute roughly
+in half.
 
@@
 for which truncation error is negligibly small.  In practice, only a
-fixed maximal truncation order $\overline M$ is accessible with
+fixed maximal truncation order $M_{\t{max}}$ is accessible with
 limited computational resources, such that
 Eq.~\eqref{eq:order_truncation} only allows computation of the
-correlator $\bk{\S_{\v m}}$ to a maximal time $t_{\overline M}$.  The
-maximal accessible truncation order $\overline M$ generally depends on
+correlator $\bk{\S_{\v m}}$ to a maximal time $t_{M_{\t{max}}}$.  The
+maximal accessible truncation order $M_{\t{max}}$ generally depends on
 the choice of operator basis for collective spin systems, which we
-discuss briefly at the end of Appendix \ref{sec:new_basis}.  As we
-will see in the following sections, our choice of basis elements
-$\S_{\v m}=S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$ allows us to accurately
-compute spin correlators to times that are relevant e.g.~for spin
-squeezing protocols.
+discuss briefly in Section \ref{sec:breakdown} and Appendix
+\ref{sec:new_basis}.  As we will see in the following sections, our
+choice of basis elements $\S_{\v m}=S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$
+allows us to accurately compute spin correlators to times that are
+relevant e.g.~for spin squeezing protocols.
 
@@
 where we include a factor of $1/3$ in the TAT Hamiltonian because it
-naturally appears in realistic proposals to experimentally realize
+naturally appears in realistic proposals to experimentally implement
 TAT\cite{liu2011spin, huang2015twoaxis}.  For simplicity, we further
@@
 obtainable with the collective-spin quantum trajectory Monte Carlo
-method recently developed in ref.~\cite{zhang2018montecarlo}.  In this
-work, these exact and quantum trajectory simulations will be used to
+method developed in ref.~\cite{zhang2018montecarlo}.  In this work,
+these exact and quantum trajectory simulations will be used to
 benchmark the expansion in Eq.~\eqref{eq:order_truncation}.
@@
   \centering
-  \includegraphics{coherent.pdf}
+  \subfloat[Squeezing under coherent dynamics \label{fig:coherent}]
+  {\includegraphics{coherent.pdf}} \\
+  \subfloat[Squeezing with decoherence:
+  $\gamma_-=\gamma_+=\gamma_\z=\chi$ \label{fig:incoherent}]
+  {\includegraphics{decoherence_weak.pdf}}
   \caption{Spin squeezing of $N=10^4$ spins initially in $\ket\X$
-    under coherent evolution, as computed using exact methods (solid
-    lines) and the expansion in Eq.~\eqref{eq:order_truncation} with
-    $M=40$ (dashed lines).  Solid circles mark the times at which the
-    latter method gives an unphysical result with $\xi^2<0$.  Both
-    methods agree exactly in the OAT case for all times shown.}
-  \label{fig:coherent}
+    under \sref{fig:coherent} coherent and \sref{fig:incoherent}
+    incoherent evolution, as computed using exact methods (solid
+    lines), quantum trajectory simulations (dots), and the expansion
+    in Eq.~\eqref{eq:order_truncation} with $M=40$ (darker dashed
+    lines).  Solid circles mark the times at which the last of these
+    methods gives an unphysical result with $\xi^2<0$.}
+  \label{fig:benchmarking}
 \end{figure}
 
-\begin{figure}
-  \centering
-  % \includegraphics{weak_decoherence.pdf}
-  \caption{Spin squeezing of $N=10^4$ spins initially in $\ket\X$ with
-    decoherence via uncorrelated decay, excitation, and dephasing of
-    individual spins at rates $\gamma_-=\gamma_+=\gamma_\z=\chi$.
-    Squeezing is computed using an exact simulation (solid line),
-    quantum trajectory methods (dots), and the expansion in
-    Eq.~\eqref{eq:order_truncation} with $M=40$ (dashed lines).  Solid
-    circles mark the times at which the latter method gives an
-    unphysical result with $\xi^2<0$.}
-  \label{fig:weak_decoherence}
-\end{figure}
-
-Figure \ref{fig:coherent} compares the squeezing parameter $\xi^2$ for
-$N=10^4$ spins initially in the state $\ket{\X}$ coherently evolved
-under the Hamiltonians in Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT}, as
-computed via both benchmarking simulations and the prescription in
-Eq.~\eqref{eq:order_truncation} with $M=40$.  Similarly, Figure
-\ref{fig:weak_decoherence} compares squeezing in the same scenario
-with the addition of decoherence through uncorrelated emission,
-absorption, and dephasing of individual spins at a rate $\chi$,
-respectively described by the sets of jump operators
-$\J_n=\set{s_n^{(j)}}$ with corresponding decoherence rates
+Figure \ref{fig:benchmarking} compares the squeezing parameter $\xi^2$
+for $N=10^4$ spins initially in the state $\ket{\X}$ evolved under the
+Hamiltonians in Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT}, as computed via
+both benchmarking simulations and the prescription in
+Eq.~\eqref{eq:order_truncation} with $M=40$.  This squeezing is shown
+for both coherent dynamics (Figure \ref{fig:coherent}), as well as
+dynamics in the presence of uncorrelated emission, absorption, and
+dephasing of individual spins at a rate $\chi$ (Figure
+\ref{fig:incoherent}), respectively described by the sets of jump
+operators $\J_n=\set{s_n^{(j)}}$ with corresponding decoherence rates
 $\gamma_n=\chi$ for $n\in\set{-,+,\z}$.  The squeezing results shown
-in Figures \ref{fig:coherent} and \ref{fig:weak_decoherence} were
-computed in a rotated basis with $\p{s_\z,s_\x}\to\p{s_\x,-s_\z}$ and
+in Figure \ref{fig:benchmarking} were computed in a rotated basis with
+$\p{s_\z,s_\x}\to\p{s_\x,-s_\z}$ and
 $\ket\X\to\ket{-\Z}\equiv\ket\dn^{\otimes N}$; the only effect of this
-rotation on the results presented in these figures is to roughly
-double the time span of near-perfect agreement in squeezing via TNT.
-The reason for different results in rotated bases has to do with the
-breakdown of the expansion in Eq.~\eqref{eq:order_truncation}, which
-we discuss further in Section \ref{sec:breakdown}.
+rotation on the results presented in this figure is to roughly double
+the time for which all TNT results agree.  The reason for different
+results in rotated bases has to do with the breakdown of the expansion
+in Eq.~\eqref{eq:order_truncation}, which we discuss further in
+Section \ref{sec:breakdown}.
 
@@
 
-\section{Many-body correlators}
-
-Many-body correlators and cumulants \`a la
-Ref.~[\citenum{schweigler2017solving}]?
+\section{Out-of-time-ordered correlators}
+
+Compare ``true'' OTOCs with decoherence vs.~``practical'' OTOCs in experiments with time-reversal of coherent dynamics and decoherence
+
+Discuss extension of this work to spin-boson systems
 
@@
   $s_\mu^{(j)}$ operators on the right throughout these calculations,
-  due to the enhanced symmetry of expressions with respect to
+  due to the enhanced symmetry expressions would have with respect to
   Hermitian conjugation.  In any case, we express the final result of
@@
 managed to order all spin operators, as $C_{m-k-1,n-1;\mu}$ contains
-$S_\z$ operators which are to the left of $S_\mu^k$.  To sort all spin
+$S_\z$ operators that are to the left of $S_\mu^k$.  To sort all spin
 operators once and for all, we define
@@
 commutator $\sp{S_\mu^m,S_\bmu^n}_- = C_{mn;\mu}$, but these result
-are -- to put it lightly -- quite a mess: the expression for
+are -- put lightly -- quite a mess: the expression for
 $E_{mn;\mu}^{(p)}$ in \eqref{eq:E_mn_p} involves a sum over $p$
@@
 
-In this section we derive several identities which will be necessary
+In this section we derive several identities that will be necessary
 for computing the effects of single-spin decoherence on ordered
@@
 \begin{align}
-  \ket{\tilde\phi} \equiv e^{-i\phi S_\z}\ket{\X},
-  &&
-  \p{\cos\phi S_\x + \sin\phi S_\y} \ket{\tilde\phi}
-  = S \ket{\tilde\phi},
+  \ket{\phi_{\x\y}} \equiv e^{-i\phi S_\z}\ket{\X},
+  &&
+  \p{\cos\phi S_\x + \sin\phi S_\y} \ket{\phi_{\x\y}}
+  = S \ket{\phi_{\x\y}},
 \end{align}
@@
 \begin{align}
-  \bk{\tilde\phi|S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n|\tilde\phi}
+  \bk{\phi_{\x\y}|S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n|\phi_{\x\y}}
   = e^{i\mu\p{\ell-n}\phi} \bk{\X|S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n|\X}.
@@
 
-\section{Changing bases for collective spin operators}
+\section{Changing operator bases, and a general product for algebras
+  of permutationally-symmetric operators}
 \label{sec:new_basis}
@@
 
-We now expand a product of basis operators $\tilde\S_{\v m}$ and
-$\tilde\S_{\v n}$ into terms that have exactly $s$ spins addressed by
-both operators:
+In order to make use of the operators $\tilde S_{\v m}$, we will need
+to compute products of the form $\tilde S_{\v m} \tilde S_{\v n}$.  We
+therefore expand such a product into terms that have exactly $s$ spins
+addressed by both operators:
 \begin{align}
@@
 $r_{\mu\nu}=\abs{\set{K^{\mu\nu}_a}}$ for brevity; and
-$\eta_{\mu\nu\rho}$ is a structure factor defined by
+$\eta_{\mu\nu\rho}$ is a structure constant defined by
 $s_\mu s_\nu=\sum_\rho\eta_{\mu\nu\rho}s_\rho$.  Unlike the sums over
@@
   \eta_{\mu\nu\rho^{\mu\nu}_a} s_{\rho^{\mu\nu}_a}
-  = \sum_{\v\rho} \sp{\prod_{\mu,\nu,a}
-    \p{\eta_{\mu\nu\rho^{\mu\nu}_a}}^{\rho^{\mu\nu}_a}}
+  = \sum_{\v\rho} \sp{\prod_{\mu,\nu,a}\eta_{\mu\nu\rho^{\mu\nu}_a}}
   \bigotimes_{\mu,\nu,a} s_{\rho^{\mu\nu}_a},
@@
 are equal up to a permutation of indices if
-$\tilde{\v\rho}_1=\tilde{\v\rho}_2$.  The degeneracy of terms in
-\eqref{eq:Q_K_rho}, i.e.~the number of $\v\rho$ which are consistent
-with $\tilde{\v\rho}$, is
+$\tilde{\v\rho}_1=\tilde{\v\rho}_2$.  The degeneracy (under
+permutation) of terms in \eqref{eq:Q_K_rho}, i.e.~the number of
+$\v\rho$ that are consistent with $\tilde{\v\rho}$, is
 \begin{align}
@@
     \choose \tilde\rho^{\mu\nu}_\kappa }
-  = \prod_{\mu,\nu}
-  \f{r_{\mu\nu}!}{\prod_\kappa\tilde\rho^{\mu\nu}_\kappa!}.
+  = \sp{\prod_{\mu,\nu} r_{\mu\nu}!}
+  \sp{\prod_{\mu,\nu,\kappa} \tilde\rho^{\mu\nu}_\kappa!}^{-1}.
 \end{align}
@@
 \begin{align}
-  \v\eta_{\tilde{\v\rho}}
+  \v\eta^{\tilde{\v\rho}}
   \equiv \prod_{\mu,\nu,\kappa}
@@
   Q_{\v K}
-  \simeq \sum_{\tilde{\v\rho}} g_{\tilde{\v\rho}} \v\eta_{\tilde{\v\rho}}
+  \simeq \sum_{\tilde{\v\rho}} g_{\tilde{\v\rho}} \v\eta^{\tilde{\v\rho}}
   \bigotimes_{\kappa}
   s_\kappa^{\otimes\sum_{\mu,\nu}\tilde\rho^{\mu\nu}_\kappa}
-  \simeq \sum_{\tilde{\v\rho}} g_{\tilde{\v\rho}} \v\eta_{\tilde{\v\rho}}
+  \simeq \sum_{\tilde{\v\rho}} g_{\tilde{\v\rho}} \v\eta^{\tilde{\v\rho}}
   \1^{\otimes\sum_{\mu,\nu}\tilde\rho^{\mu\nu}_0}
@@
   = \sum_{s\ge0} \sum_{\p{\v r;\v m,\v n,s}} f_{\v m\v n\v r}
-  \sum_{\p{\v\rho;\v r}} g_{\v\rho} \v\eta_{\v\rho}
+  \sum_{\p{\v\rho;\v r}} g_{\v\rho} \v\eta^{\v\rho}
   c_{\v\ell_{\v m\v n\v r\v\rho}\v\rho}
@@
 \end{align}
-accounts for the number of terms which are equivalent up to a
+accounts for the number of terms that are equivalent up to a
 permutation of indices involving the identity operators in
@@
 can, however, impose additional restrictions on these sums in such a
-way as to throw out terms which vanish, and keep only those which do
-not.  Remembering that $s$ counts the number of spins which are
+way as to throw out all terms that are zero, and keep only those that
+are not.  Remembering that $s$ counts the number of spins that are
 addressed by both operators $P_{\v j}$ and $P_{\v k}$ in
@@
 \end{align}
-In addition, the structure factor $\eta_{\mu\nu\kappa}$ may generally
-contain zeros that we can preemptively avoid with restrictions on
-$\v r$ and $\v\rho$.  Recalling that $r_{\mu\nu}$ counts the number of
-operators addressed by
+In addition, the matrix of structure constants $\eta_{\mu\nu\kappa}$
+may generally contain zeros that we can preemptively avoid with
+restrictions on $\v r$ and $\v\rho$.  Recalling that $r_{\mu\nu}$
+counts the number of operators addressed by
 $s_\mu s_\nu=\sum_\kappa\eta_{\mu\nu\kappa} s_\kappa$, if
@@
 all $\mu,\nu$ with $s_\mu s_\nu=0$.  Finally, we can avoid the
-nullifying effect of zeros from
-$\v\eta_{\v\rho}=\prod_{\mu,\nu,\kappa}
+nullifying effect of zeros from the factor
+$\v\eta^{\v\rho}=\prod_{\mu,\nu,\kappa}
 \p{\eta_{\mu\nu\kappa}}^{\rho^{\mu\nu}_\kappa}$ by fixing

### 2019-03-18 multi_body_interactions/papers/theory/paper.tex
final fixes/revisions to multi-body theory paper

@@
 % "left vector" arrow; requires tikz package
+\usepackage{tikz}
 \newcommand{\lvec}[1]
@@
 
-While an external trapping potential will generally break discrete
+Although an external trapping potential will generally break discrete
 translational symmetry of the lattice, any background inhomogeneity
@@
 atomic density field operator; and the coupling constants $G_X$ are
-defined in terms of the scattering lengths by
+defined in terms of the scattering lengths $a_Y$ by
 \begin{align}
@@
   \c_{\nu s}^\dag \c_{\mu t}^\dag \c_{\nu r} \c_{\mu q}
-  = -G^{qr}_{ts} \c_{\nu s}^\dag \c_{\mu t}^\dag \c_{\mu r} \c_{\nu q}
+  = -G^{qr}_{ts} \c_{\nu s}^\dag \c_{\mu t}^\dag \c_{\nu r} \c_{\mu q}
   = G^{qr}_{ts} \c_{\mu t}^\dag \c_{\nu s}^\dag \c_{\nu r} \c_{\mu q}.
@@
 \begin{align}
-  \ket{\N,-_j} \equiv \f1{\sqrt2}
+  \ket{\N,-,j} \equiv \f1{\sqrt2}
   \p{\c_{\mu_1,\e}^\dag \c_{\mu_1,\g} - \c_{\mu_j,\e}^\dag \c_{\mu_j,\g}}

### 2019-03-20 soc_squeezing/collective_paper/collective_spin.tex
added discussion of advantages and limitations of the truncated Taylor expansion method

@@
 Collective spin systems are a versatile resource in quantum science
-for a range of applications including quantum-enhanced metrology,
-quantum simulation, and quantum information.  Theoretical interest in
-such systems dates back to the mid-twentieth century with the
-introduction of the Dicke model\cite{dicke1954coherence} that
-describes atoms cooperatively interacting with a single mode of the
-radiation field and the Lipkin-Meshkov-Glick (LMG) model, a toy model
-for testing many-body approximation methods in contemporary nuclear
+for a range of applications including quantum-enhanced metrology and
+quantum simulation.  Theoretical interest in such systems dates back
+to the mid-twentieth century with the introduction of the Dicke
+model\cite{dicke1954coherence} that describes atoms cooperatively
+interacting with a single mode of a radiation field, and the
+Lipkin-Meshkov-Glick (LMG) model, a toy model for testing many-body
+approximation methods in contemporary nuclear
 physics\cite{lipkin1965validity, meshkov1965validity,
@@
   hosten2016quantum, lewis-swan2018robust, norcia2018cavitymediated,
-  jones2009magnetic}.
+  jones2009magnetic, he2019engineering}.
 
@@
   hosten2016quantum, lewis-swan2018robust, norcia2018cavitymediated,
-  ma2011quantum} that allow for measurement errors
+  ma2011quantum, he2019engineering} that allow for measurement errors
 $\Delta\phi\sim1/N^\epsilon$ with $\epsilon>1/2$.  Generating such
@@
 interactions, as long as measurements do not distinguish between
-constituent particles even non-uniform systems can be effectively
+constituent particles, even non-uniform systems can be effectively
 described by a uniform model with renormalized
@@
 jumps that generate large numbers of distinct quantum trajectories
-that need to be averaged to compute expectation values.
-
-In this work, we develop a new method to simulate short-time dynamics
+that need to be averaged over to compute expectation values.
+
+In this work, we present a new method to simulate short-time dynamics
 of collective spin systems with permutationally-symmetric decoherence.
@@
 expansion requires using the structure constants of a collective spin
-operator algebra; the calculation of these structure constants is one
-of the main technical results of this work, which we hope will enable
-future studies to develop new analytical insights into the dynamics of
-collective spin systems.
-
-[Paragraph or so about benchmarking and the discussion section?]
+operator algebra; the calculation of these structure constants (in
+Appendices \ref{sec:identities}--\ref{sec:prod_general}) is one of the
+main technical results of this work, which we hope will enable future
+studies to develop new analytical insights into the behavior of
+collective spin systems.  We benchmark our method against both exact
+and quantum trajectory Monte Carlo calculations of spin squeezing in
+accessible parameter regimes, highlighting both advantages and
+limitations of the short-time expansion.  Finally, we showcase an
+application of our method with calculations that are inaccessible to
+other methods.
+
+[More text about the content of this paper.]
 
@@
 expectation values of collective spin operators, deferring lengthy
-derivations to the Appendices.  We consider systems composed of $N$
+derivations to the appendices.  We consider systems composed of $N$
 distinct spin-1/2 particles.  Defining individual spin-1/2 operators
@@
 individual spins at a rate $\gamma_-$, for example, would be described
-by the set of jump operators
-$\J_-=\set{\sigma_-^{(j)}:j=1,2,\cdots,N}$ with
-$\gamma_{\J_-}=\gamma_-$.  The commutator in Eq.~\eqref{eq:EOM} can be
-determined from the structure constants $f_{\v n\v m\v\ell}$ of the
-general product
-$\S_{\v n}\S_{\v m}=\sum_{\v\ell}f_{\v n\v m\v\ell}\S_{\v\ell}$ worked
+by the set of jump operators $\J_-=\set{s_-^{(j)}:j=1,2,\cdots,N}$
+with $\gamma_{\J_-}=\gamma_-$.  The commutator in Eq.~\eqref{eq:EOM}
+can be determined from the structure constants $f_{\v\ell\v m\v n}$ of
+the general product
+$\S_{\v\ell}\S_{\v m}=\sum_{\v n}f_{\v\ell\v m\v n}\S_{\v n}$ worked
 out in Appendices \ref{sec:identities}--\ref{sec:prod_general}, while
@@
 of the form $\gamma^{(j)}=\sum_n\gamma_ns_n^{(j)}$ and
-$\Gamma=\sum_n\Gamma_nS_n$ are respectively worked out in Appendices
-\ref{sec:sandwich_single}--\ref{sec:decoherence_collective}.  These
-ingredients are sufficient to compute matrix elements $T_{\v m\v n}$
-of the time derivative operator $d/dt$ in Eq.~\eqref{eq:EOM} in most
-cases of experimental relevance.
+$\Gamma=\sum_n\Gamma_nS_n$ are worked out in Appendices
+\ref{sec:sandwich_single}--\ref{sec:decoherence_collective}.  We
+consider these calculations one of the main technical contributions of
+our work, with potential applications beyond the short-time simulation
+method presented here.  These ingredients are sufficient to compute
+matrix elements $T_{\v m\v n}$ of the time derivative operator $d/dt$
+in Eq.~\eqref{eq:EOM} in most cases of practical interest.
 
@@
 
-In principle, the factorial suppression of terms at higher orders
-(i.e.~in $k$) of the expansion in Eq.~\eqref{eq:order_truncation}
-implies for any given time $t$, there exists a truncation order $M_t$
-for which truncation error is negligibly small.  In practice, only a
-fixed maximal truncation order $M_{\t{max}}$ is accessible with
-limited computational resources, such that
-Eq.~\eqref{eq:order_truncation} only allows computation of the
-correlator $\bk{\S_{\v m}}$ to a maximal time $t_{M_{\t{max}}}$.  The
-maximal accessible truncation order $M_{\t{max}}$ generally depends on
-the choice of operator basis for collective spin systems, which we
-discuss briefly in Section \ref{sec:breakdown} and Appendix
-\ref{sec:new_basis}.  As we will see in the following sections, our
-choice of basis elements $\S_{\v m}=S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$
-allows us to accurately compute spin correlators to times that are
-relevant e.g.~for spin squeezing protocols.
-
-\section{Spin squeezing}
+In principle, the factorial suppression of terms at higher orders of
+the Taylor expansion in Eq.~\eqref{eq:time_series} implies for any
+given time $t$, there exists a truncation order $M_t$ for which the
+truncation error in Eq.~\eqref{eq:order_truncation} is negligibly
+small.  In practice, only a maximal truncation order $M_{\t{max}}$ is
+accessible with limited computational resources, such that the
+truncated Taylor expansion in Eq.~\eqref{eq:order_truncation} only
+allows computation of the correlator $\bk{\S_{\v m}}$ to a maximal
+time $t_{M_{\t{max}}}$.  As we will see in the following section,
+$M_{\t{max}}=40$ will suffice for the accurate computation of
+collective spin correlators up to times that are relevant e.g.~for
+spin squeezing protocols.
+
+\section{Spin squeezing, benchmarking, and breakdown}
 
@@
 \end{align}
-where $\v S\equiv\p{S_\z,S_\x,S_\y}$ is a collective spin vector, and
+where $\v S\equiv\p{S_\x,S_\y,S_\z}$ is a collective spin vector, and
 the minimization is performed over all unit vectors $\uv n$ in the
@@
 these exact and quantum trajectory simulations will be used to
-benchmark the expansion in Eq.~\eqref{eq:order_truncation}.
+benchmark the truncated Taylor expansion in
+Eq.~\eqref{eq:order_truncation}.
 
@@
   \centering
-  \subfloat[Squeezing under coherent dynamics \label{fig:coherent}]
+  \subfloat[Squeezing under coherent dynamics
+  \label{fig:squeezing_coherent}]
   {\includegraphics{coherent.pdf}} \\
   \subfloat[Squeezing with decoherence:
-  $\gamma_-=\gamma_+=\gamma_\z=\chi$ \label{fig:incoherent}]
+  $\gamma_-=\gamma_+=\gamma_\z=\chi$
+  \label{fig:squeezing_incoherent}]
   {\includegraphics{decoherence_weak.pdf}}
   \caption{Spin squeezing of $N=10^4$ spins initially in $\ket\X$
-    under \sref{fig:coherent} coherent and \sref{fig:incoherent}
-    incoherent evolution, as computed using exact methods (solid
-    lines), quantum trajectory simulations (dots), and the expansion
-    in Eq.~\eqref{eq:order_truncation} with $M=40$ (darker dashed
-    lines).  Solid circles mark the times at which the last of these
-    methods gives an unphysical result with $\xi^2<0$.}
+    under \sref{fig:squeezing_coherent} coherent and
+    \sref{fig:squeezing_incoherent} incoherent evolution, computed
+    using exact methods (solid lines), quantum trajectory simulations
+    (dots), and the truncated Taylor expansion in
+    Eq.~\eqref{eq:order_truncation} with $M=40$ (darker dashed lines).
+    Solid circles mark the times at which the truncated Taylor
+    expansion gives an unphysical result with $\xi^2<0$.}
   \label{fig:benchmarking}
@@
 Hamiltonians in Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT}, as computed via
-both benchmarking simulations and the prescription in
-Eq.~\eqref{eq:order_truncation} with $M=40$.  This squeezing is shown
-for both coherent dynamics (Figure \ref{fig:coherent}), as well as
-dynamics in the presence of uncorrelated emission, absorption, and
+both benchmarking simulations and the truncated Taylor expansion in
+Eq.~\eqref{eq:order_truncation} with $M=40$.  Squeezing is shown for
+both coherent dynamics (Figure \ref{fig:squeezing_coherent}), as well
+as dynamics in the presence of uncorrelated emission, absorption, and
 dephasing of individual spins at a rate $\chi$ (Figure
-\ref{fig:incoherent}), respectively described by the sets of jump
-operators $\J_n=\set{s_n^{(j)}}$ with corresponding decoherence rates
-$\gamma_n=\chi$ for $n\in\set{-,+,\z}$.  The squeezing results shown
-in Figure \ref{fig:benchmarking} were computed in a rotated basis with
+\ref{fig:squeezing_incoherent}), respectively described by the sets of
+jump operators $\J_n=\set{s_n^{(j)}}$ with corresponding decoherence
+rates $\gamma_n=\chi$ for $n\in\set{-,+,\z}$.  The results shown in
+Figure \ref{fig:benchmarking} were computed in a rotated basis with
 $\p{s_\z,s_\x}\to\p{s_\x,-s_\z}$ and
-$\ket\X\to\ket{-\Z}\equiv\ket\dn^{\otimes N}$; the only effect of this
-rotation on the results presented in this figure is to roughly double
-the time for which all TNT results agree.  The reason for different
-results in rotated bases has to do with the breakdown of the expansion
-in Eq.~\eqref{eq:order_truncation}, which we discuss further in
-Section \ref{sec:breakdown}.
-
-[Comment on near-perfect agreement right up until a sudden and drastic
-departure from the correct result.  Reference figure showing squeezing
-for large system with strong decoherence.  Mention that this method
-was used to put a lower bound on squeezing via TAT in our
-SOC+interactions squeezing paper.]
+$\ket\X\to\ket{-\Z}\equiv\ket\dn^{\otimes N}$ (together with
+appropriate transformations of the Hamiltonian and jump operators);
+the only effect of this rotation on the results presented in Figure
+\ref{fig:benchmarking} is to prolong the time for which the truncated
+Taylor expansion agrees with the benchmarking simulations.  The reason
+for different results in rotated basis has to do with the breakdown of
+the truncated Taylor expansion, which we discuss further below.
+
+The main lesson from Figure \ref{fig:benchmarking} is that the
+truncated Taylor expansion in Eq.~\eqref{eq:order_truncation} gives
+essentially exact results right up until a sudden and drastic
+departure that is simple to diagnose by inspection.  This drastic
+departure occurs because the truncated Taylor expansion neglects
+high-weight (i.e.~large-$\abs{\v n}$) operators $\S_{\v n}$ whose
+contributions to a correlator of interest $\bk{\S_{\v m}}$ eventually
+become non-negligible.  For sufficiently large times $t$, these
+individual contributions may generally be large compared to the value
+of $\bk{\S_{\v m}}$, in which case their truncation will yield
+non-physical results.  This breakdown mechanism is also the reason for
+prolonged agreement between the truncated Taylor expansion and
+benchmarking methods when squeezing the initial state $\ket{-\Z}$
+rather than $\ket{\X}$: for initial states $\ket{-\Z}$, all
+correlators $\bk{\S_{\v n}}_{t=0}$ vanish unless $n_+=n_-=0$ (see
+Appendix \ref{sec:initial_conditions}), so there is a substantially
+smaller number of terms with non-zero contribution to the correlator
+$\bk{\S_{\v m}}$, and therefore a substantially smaller number of
+non-zero terms that are neglected in the truncated Taylor expansion.
+
+Although the truncated Taylor expansion breaks down at short times, it
+has two key advantages over other methods to compute collective spin
+correlators.  First, computing spin correlators with the truncated
+Taylor expansion is generally much faster and requires much fewer
+computing resources than alternatives.  The quantum trajectory Monte
+Carlo simulations performed for Figure \ref{fig:squeezing_incoherent},
+for example, take $\sim10^4$ CPU hours to compute on standard modern
+computing hardware; the bulk of this time is spent performing sparse
+matrix multiplication, leaving little room to further optimize run
+time.  Parallelization can reduce the actual run time of these quantum
+trajectory simulations to $\sim10$ hours, but at the cost of greatly
+increasing computing resource requirements.  The truncated Taylor
+expansion results in Figure \ref{fig:squeezing_incoherent}, meanwhile,
+take $\sim10$ seconds to compute with a single CPU on similar
+computing hardware.
+
+Second, the truncated Taylor expansion enables computing spin
+correlators in strong-decoherence regimes entirely inaccessible to
+other methods.  As an example, Figure \ref{fig:decoherence_strong}
+shows squeezing of $N=10^4$ spins initially in $\ket{\X}$, with decay
+and dephasing of individual spins at rates
+$\gamma_-=\gamma_\z=100\chi$.  Computations of this sort were used to
+put lower bounds on theoretically achievable spin squeezing via TAT
+with decoherence in Ref.~[\citenum{he2019engineering}], exemplifying a
+concrete application of the truncated Taylor expansion and the
+collective-spin structure constants calculated in this work.
+
+\begin{figure}
+  \centering
+  \includegraphics{decoherence_strong.pdf}
+  \caption{Spin squeezing of $N=10^4$ spins initially in $\ket\X$
+    under incoherent evolution with decay and dephasing of individual
+    spins at rates $\gamma_-=\gamma_\z=100\chi$, computed using the
+    truncated Taylor expansion in Eq.~\eqref{eq:order_truncation} with
+    $M=40$.  Solid circles mark the times at which the truncated
+    Taylor expansion gives an unphysical result with $\xi^2<0$.}
+  \label{fig:decoherence_strong}
+\end{figure}
 
 \section{Out-of-time-ordered correlators}
-
-Compare ``true'' OTOCs with decoherence vs.~``practical'' OTOCs in experiments with time-reversal of coherent dynamics and decoherence
+\label{sec:OTOC}
+
+Compare ``true'' OTOCs with decoherence vs.~``practical'' OTOCs in
+experiments with time-reversal of coherent dynamics and decoherence
 
 Discuss extension of this work to spin-boson systems
-
-\section{Breakdown}
-\label{sec:breakdown}
 
@@
   due to the enhanced symmetry expressions would have with respect to
-  Hermitian conjugation.  In any case, we express the final result of
-  this section in both ordering conventions, and thus feel no need to
-  reproduce these calculations with a different ordering.}.  This
-choice of ordering compels us to expand
+  Hermitian conjugation.  In any case, we provide the final result of
+  this section in both ordering conventions, and therefore feel no
+  need to reproduce these calculations with a different ordering of
+  spin operators.}.  This choice of ordering compels us to expand
 \begin{align}

### 2019-03-21 soc_squeezing/collective_paper/collective_spin.tex
updated text to reflect changes to figures in plot_squeezing.py

@@
 time $t_{M_{\t{max}}}$.  As we will see in the following section,
-$M_{\t{max}}=40$ will suffice for the accurate computation of
+$M_{\t{max}}=35$ will suffice for the accurate computation of
 collective spin correlators up to times that are relevant e.g.~for
@@
     (dots), and the truncated Taylor expansion in
-    Eq.~\eqref{eq:order_truncation} with $M=40$ (darker dashed lines).
+    Eq.~\eqref{eq:order_truncation} with $M=35$ (darker dashed lines).
     Solid circles mark the times at which the truncated Taylor
@@
 both benchmarking simulations and the truncated Taylor expansion in
-Eq.~\eqref{eq:order_truncation} with $M=40$.  Squeezing is shown for
+Eq.~\eqref{eq:order_truncation} with $M=35$.  Squeezing is shown for
 both coherent dynamics (Figure \ref{fig:squeezing_coherent}), as well
-as dynamics in the presence of uncorrelated emission, absorption, and
-dephasing of individual spins at a rate $\chi$ (Figure
+as dynamics in the presence of spontaneous decay, excitation, and
+dephasing of individual spins at rates $\chi$ (Figure
 \ref{fig:squeezing_incoherent}), respectively described by the sets of
@@
 other methods.  As an example, Figure \ref{fig:decoherence_strong}
-shows squeezing of $N=10^4$ spins initially in $\ket{\X}$, with decay
-and dephasing of individual spins at rates
-$\gamma_-=\gamma_\z=100\chi$.  Computations of this sort were used to
-put lower bounds on theoretically achievable spin squeezing via TAT
-with decoherence in Ref.~[\citenum{he2019engineering}], exemplifying a
-concrete application of the truncated Taylor expansion and the
-collective-spin structure constants calculated in this work.
+shows squeezing of $N=10^4$ spins initially in $\ket{\X}$, undergoing
+spontaneous decay, excitation, and dephasing of individual spins at
+rates $\gamma_-=\gamma_+=\gamma_\z=100\chi$.  Computations of this
+sort were used to put lower bounds on theoretically achievable spin
+squeezing via TAT with decoherence in
+Ref.~[\citenum{he2019engineering}], exemplifying a concrete
+application of the truncated Taylor expansion and the collective-spin
+structure constants calculated in this work.
 
@@
   \caption{Spin squeezing of $N=10^4$ spins initially in $\ket\X$
-    under incoherent evolution with decay and dephasing of individual
-    spins at rates $\gamma_-=\gamma_\z=100\chi$, computed using the
+    under incoherent evolution spontaneous with decay, excitation, and
+    dephasing of individual spins at rates
+    $\gamma_-=\gamma_+=\gamma_\z=100\chi$, computed using the
     truncated Taylor expansion in Eq.~\eqref{eq:order_truncation} with
-    $M=40$.  Solid circles mark the times at which the truncated
+    $M=35$.  Solid circles mark the times at which the truncated
     Taylor expansion gives an unphysical result with $\xi^2<0$.}
@@
 
-\section{Out-of-time-ordered correlators}
-\label{sec:OTOC}
-
-Compare ``true'' OTOCs with decoherence vs.~``practical'' OTOCs in
-experiments with time-reversal of coherent dynamics and decoherence
-
-Discuss extension of this work to spin-boson systems
+\section{Two-time correlation functions}
+\label{sec:TTC}
+
+These are very easy to compute.
+
+% \section{Out-of-time-ordered correlators}
+% \label{sec:OTOC}
+
+% Compute some OTOCs?  There is a small chance that doing this will
+% take a prohibitively long amount of time...
+
+% Possibility: compare ``true'' OTOCs with decoherence
+% vs.~``practical'' OTOCs in experiments with decoherence and
+% time-reversal of coherent dynamics.
 
 \section{Conclusions}
+
+Mention/discuss extension of this work to spin-boson systems.
 

### 2019-03-26 soc_squeezing/collective_paper/collective_spin.tex
wording changes and minor typo fixes

@@
 Appendices \ref{sec:identities}--\ref{sec:prod_general}) is one of the
-main technical results of this work, which we hope will enable future
-studies to develop new analytical insights into the behavior of
-collective spin systems.  We benchmark our method against both exact
-and quantum trajectory Monte Carlo calculations of spin squeezing in
-accessible parameter regimes, highlighting both advantages and
-limitations of the short-time expansion.  Finally, we showcase an
-application of our method with calculations that are inaccessible to
-other methods.
+main technical results of this work, which we hope will empower both
+analytical and numerical future studies of collective spin systems.
+We benchmark our method against both exact and quantum trajectory
+Monte Carlo computations of spin squeezing in accessible parameter
+regimes, highlighting both advantages and limitations of our
+short-time expansion.  Finally, we showcase an application of our
+method with computations that are inaccessible to other numerical
+methods.
 
@@
 \ref{sec:sandwich_single}--\ref{sec:decoherence_collective}.  We
-consider these calculations one of the main technical contributions of
+consider these calculations to be the main technical contributions of
 our work, with potential applications beyond the short-time simulation
@@
 
-We can expand Heisenberg operators in a Taylor series about $t=0$ to
-write
-\begin{align}
-  \bk{\S_{\v m}}
-  = \sum_{k\ge0} \f{t^k}{k!} \bk{\f{d^k}{dt^k} \S_{\v m}}_{t=0}
-  = \sum_{k\ge0} \f{t^k}{k!}
-  \sum_{\v n} T^k_{\v m\v n} \bk{\S_{\v n}}_{t=0},
+We can expand Heisenberg operators in a Taylor series about time $t=0$
+to write
+\begin{align}
+  \S_{\v m}\p{t}
+  % = \left. \exp\p{t \f{d}{d\tau}} \S_{\v m}\p{\tau} \right|_{\tau=0}
+  &= \sum_{k\ge0} \f{t^k}{k!}
+  \left. \f{d^k}{d\tau^k} \S_{\v m}\p{\tau} \right|_{\tau=0} \\
+  &= \sum_{k\ge0} \f{t^k}{k!}
+  \sum_{\v n} T^k_{\v m\v n} \S_{\v n}\p{0},
   \label{eq:time_series}
@@
 \begin{align}
-  \bk{\S_{\v m}}
+  \S_{\v m}\p{t}
   \to \sum_{k=0}^M \f{t^k}{k!}
-  \sum_{\v n} T^k_{\v m\v n} \bk{\S_{\v n}}_{t=0}.
+  \sum_{\v n} T^k_{\v m\v n} \S_{\v n}\p{0}.
   \label{eq:order_truncation}
@@
 which operators are relevant for computing the expectation value
-$\bk{\S_{\v m}}$ to a fixed order $M$.  Note that the relation
+$\bk{\S_{\v m}\p{t}}$ to a fixed order $M$.  Note that the relation
 $\S_{\v m^*}=\S_{\v m}^\dag$, which by Hermitian conjugation of
@@
 $T_{\v m^*\v n^*}=T_{\v m\v n}^*$, can be exploited to cut both the
-number of expectation values $\bk{\S_{\v m}}_{t=0}$ and matrix
-elements $T_{\v m\v n}$ that we may need to explicitly compute roughly
-in half.
+number of initial-time expectation values $\bk{\S_{\v m}\p{0}}$ and
+the number of matrix elements $T_{\v m\v n}$ that we may need to
+explicitly compute roughly in half.
 
@@
 truncated Taylor expansion in Eq.~\eqref{eq:order_truncation} only
-allows computation of the correlator $\bk{\S_{\v m}}$ to a maximal
-time $t_{M_{\t{max}}}$.  As we will see in the following section,
-$M_{\t{max}}=35$ will suffice for the accurate computation of
+allows computation of the correlator $\bk{\S_{\v m}\p{t}}$ to a
+maximal time $t_{M_{\t{max}}}$.  As we will see in the following
+section, $M_{\t{max}}=35$ will suffice for the computation of
 collective spin correlators up to times that are relevant e.g.~for
@@
 parameter is entirely determined by one- and two-spin correlators of
-the form $\bk{S_m}$ and $\bk{S_m S_n}$.  In the case of coherent
+the form $\bk{S_m}$ and $\bk{S_m S_n}$.  In the case of the coherent
 dynamics discussed here, these correlators are obtainable via exact
@@
 rather than $\ket{\X}$: for initial states $\ket{-\Z}$, all
-correlators $\bk{\S_{\v n}}_{t=0}$ vanish unless $n_+=n_-=0$ (see
-Appendix \ref{sec:initial_conditions}), so there is a substantially
-smaller number of terms with non-zero contribution to the correlator
-$\bk{\S_{\v m}}$, and therefore a substantially smaller number of
-non-zero terms that are neglected in the truncated Taylor expansion.
+initial-time correlators $\bk{\S_{\v n}\p{0}}$ vanish unless
+$n_+=n_-=0$ (see Appendix \ref{sec:initial_conditions}), so there is a
+substantially smaller number of terms with non-zero contribution to
+the correlator $\bk{\S_{\v m}}$, and therefore a substantially smaller
+number of non-zero terms that are neglected in the truncated Taylor
+expansion.
 
@@
 Taylor expansion is generally much faster and requires much fewer
-computing resources than alternatives.  The quantum trajectory Monte
-Carlo simulations performed for Figure \ref{fig:squeezing_incoherent},
-for example, take $\sim10^4$ CPU hours to compute on standard modern
-computing hardware; the bulk of this time is spent performing sparse
-matrix multiplication, leaving little room to further optimize run
-time.  Parallelization can reduce the actual run time of these quantum
-trajectory simulations to $\sim10$ hours, but at the cost of greatly
-increasing computing resource requirements.  The truncated Taylor
-expansion results in Figure \ref{fig:squeezing_incoherent}, meanwhile,
-take $\sim10$ seconds to compute with a single CPU on similar
-computing hardware.
+computing resources than the alternatives.  The quantum trajectory
+Monte Carlo simulations performed for Figure
+\ref{fig:squeezing_incoherent}, for example, take $\sim10^4$ CPU hours
+to compute on standard modern computing hardware; the bulk of this
+time is spent performing sparse matrix multiplication, leaving little
+room to further optimize run time.  Parallelization can reduce the
+actual run time of these quantum trajectory simulations to $\sim10$
+hours, but at the cost of greatly increasing computing resource
+requirements.  The truncated Taylor expansion results in Figure
+\ref{fig:squeezing_incoherent}, meanwhile, take $\sim10$ seconds to
+compute with a single CPU on similar computing hardware.
 
@@
 
-These are very easy to compute.
+As a final example, we consider the calculation of two-time
+correlation functions in collective spin systems.
 
@@
 \end{align}
-The expectation values in \eqref{eq:S+_OAT}-\eqref{eq:SS+-_OAT} and
+The expectation values in \eqref{eq:S+_OAT}--\eqref{eq:SS+-_OAT} and
 \eqref{eq:Sz_OAT} are sufficient to compute the spin squeezing
@@
 \end{align}
-In principle, the expressions in \eqref{eq:D_mn_k} and
-\eqref{eq:C_mn_E}-\eqref{eq:E_mn_p} can be used to evaluate the
-commutator $\sp{S_\mu^m,S_\bmu^n}_- = C_{mn;\mu}$, but these result
-are -- put lightly -- quite a mess: the expression for
-$E_{mn;\mu}^{(p)}$ in \eqref{eq:E_mn_p} involves a sum over $p$
-mutually dependent intermediate variables, each term of which
-additionally contains a product of $p$ factors.  We therefore devote
-the rest of this section to simplifying our result for the commutator
+In principle, the expressions in \eqref{eq:D_mn}, \eqref{eq:D_mn_k},
+\eqref{eq:C_mn_E}, and \eqref{eq:E_mn_p} suffice to evaluate the
+commutator $\sp{S_\mu^m,S_\bmu^n}_- = C_{mn;\mu}$, but this result is
+-- put lightly -- quite a mess: the expression for $E_{mn;\mu}^{(p)}$
+in \eqref{eq:E_mn_p} involves a sum over $p$ mutually dependent
+intermediate variables, each term of which additionally contains a
+product of $p$ factors.  We therefore devote the rest of this section
+to simplifying our result for the commutator
 $\sp{S_\mu^m,S_\bmu^n}_-$.
@@
   p! { m \choose p } { n \choose p } S_\bmu^{n-p} S_\mu^{m-p}
-  \sum_{\ell=0}^p \epsilon_{mn}^{p\ell} \p{\mu S_\z}^\ell,
+  \sum_{\ell=0}^p \epsilon_{mn}^{p\ell} \p{\mu S_\z}^\ell.
   \label{eq:push_mu_bmu}

### 2019-03-26 soc_squeezing/collective_paper/collective_spin.tex
changed theory section to consider the evolution of a general Heisenberg operator

@@
 $\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$, we can expand all
-collective spin Hamiltonians in the form
-\begin{align}
-  H = \sum_{\v m} h_{\v m} \S_{\v m},
-  \label{eq:H_general}
-\end{align}
-where Hermiticity requires that $h_{\v m}^*=h_{\v m^*}$ for
-$\v m^*\equiv\p{m_-,m_\z,m_+}$.  The time evolution of a particular
-Heisenberg operator $\S_{\v m}$ under a Hamiltonian of the form in
-Eq.~\eqref{eq:H_general} is then given by
-\begin{align}
-  \f{d}{dt}\S_{\v m}
-  &= i\sum_{\v n}h_{\v n}\sp{\S_{\v n},\S_{\v m}}_-
-  + \sum_\J \gamma_\J \D\p{\J;\S_{\v m}} \\
-  &= \sum_{\v n} T_{\v m\v n} \S_{\v n},
+collective spin operators in the form
+\begin{align}
+  \O = \sum_{\v m} \O_{\v m} \S_{\v m}.
+\end{align}
+The time evolution of a Heisenberg operator $\O\p{t}$ under a
+Hamiltonian $H$ is then given by
+\begin{align}
+  \f{d}{dt} \O\p{t}
+  = T \O\p{t}
+  = \sum_{\v m, \v n} \S_{\v m} T_{\v m\v n} \O_{\v n}\p{t}
   \label{eq:EOM}
 \end{align}
+for a time derivative (super-)operator $T$ with matrix elements
+$T_{\v m\v n}$, determined by
+\begin{align}
+  T \O \equiv i \sp{H, \O}_- + \sum_\J \gamma_\J \D\p{\J; \O},
+  \label{eq:time_deriv}
+\end{align}
 where $\sp{X,Y}_\pm\equiv XY\pm YX$; $\J$ is a set of jump operators
-with a corresponding decoherence rate $\gamma_\J$;
+with a corresponding decoherence rate $\gamma_\J$; and
 \begin{align}
@@
 \end{align}
-is the Heisenberg picture dissipator, or Lindblad superoperator, for
-all jump operators $J\in\J$; and $T_{\v m\v n}$ is a matrix element of
-the time derivative operator.  Decoherence via uncorrelated decay of
-individual spins at a rate $\gamma_-$, for example, would be described
-by the set of jump operators $\J_-=\set{s_-^{(j)}:j=1,2,\cdots,N}$
-with $\gamma_{\J_-}=\gamma_-$.  The commutator in Eq.~\eqref{eq:EOM}
+is the Heisenberg dissipator, or Lindblad superoperator, for all jump
+operators $J\in\J$.  Decoherence via uncorrelated decay of individual
+spins at a rate $\gamma_-$, for example, would be described by the set
+of jump operators $\J_-=\set{s_-^{(j)}:j=1,2,\cdots,N}$ with
+$\gamma_{\J_-}=\gamma_-$.  The commutator in Eq.~\eqref{eq:time_deriv}
 can be determined from the structure constants $f_{\v\ell\v m\v n}$ of
@@
 method presented here.  These ingredients are sufficient to compute
-matrix elements $T_{\v m\v n}$ of the time derivative operator $d/dt$
-in Eq.~\eqref{eq:EOM} in most cases of practical interest.
-
-The time derivative operator $d/dt$ will generally couple spin
-operators $\S_{\v m}$ to spin operators $\S_{\v n}$ with higher
-``weight'', i.e.~with $\abs{\v n}>\abs{\v m}$, where
-$\abs{\v\ell}\equiv\sum_\lambda\ell_\lambda$.  The growth of operator
+matrix elements $T_{\v m\v n}$ of the time derivative operator $T$ in
+Eq.~\eqref{eq:time_deriv} in most cases of practical interest.
+
+The time derivative operator $T$ will generally couple spin operators
+$\S_{\v n}$ (i.e.~in $\O$) to spin operators $\S_{\v m}$ with higher
+``weight'', i.e.~with $\abs{\v m}>\abs{\v n}$, where
+$\abs{\v\ell}\equiv\sum_\alpha\ell_\alpha$.  The growth of operator
 weight signifies the growth of many-body correlations.  In practice,
@@
 \begin{align}
-  \f{d}{dt} \S_{\v m}
-  \to \sum_{w\p{\v n}<W} T_{\v m\v n} \S_{\v n}
+  \f{d}{dt} \O
+  \to \sum_{w\p{\v m}<W} \S_{\v m}
+  \sum_{\v n} T_{\v m\v n} \O_{\v n}
   \label{eq:weight_truncation}
 \end{align}
-for some weight measure $w$, e.g.~$w\p{\v n}=\abs{\v n}$, and cutoff
+for some weight measure $w$, e.g.~$w\p{\v m}=\abs{\v m}$, and cutoff
 $W$.  The truncation in Eq.~\eqref{eq:weight_truncation} closes the
 system of differential equations defined by Eq.~\eqref{eq:EOM}, and
-allow us to solve it using standard numerical methods.  Initial
+allows us to solve it using standard numerical methods.  Initial
 conditions for this system of differential equations, i.e.~expectation
@@
 
-We can expand Heisenberg operators in a Taylor series about time $t=0$
-to write
-\begin{align}
-  \S_{\v m}\p{t}
-  % = \left. \exp\p{t \f{d}{d\tau}} \S_{\v m}\p{\tau} \right|_{\tau=0}
-  &= \sum_{k\ge0} \f{t^k}{k!}
-  \left. \f{d^k}{d\tau^k} \S_{\v m}\p{\tau} \right|_{\tau=0} \\
-  &= \sum_{k\ge0} \f{t^k}{k!}
-  \sum_{\v n} T^k_{\v m\v n} \S_{\v n}\p{0},
+We can expand Heisenberg operators $\O\p{t}$ in a Taylor series about
+the time $t=0$ to write
+\begin{align}
+  \O\p{t}
+  = e^{t T} \O\p{0}
+  % = \sum_{k\ge0} \f{t^k}{k!} T^k \O\p{0}
+  = \sum_{k\ge0} \f{t^k}{k!}
+  \sum_{\v m, \v n} \S_{\v m} T^k_{\v m\v n} \O_{\v n}\p{0},
   \label{eq:time_series}
@@
   &\equiv \sum_{\v p_1,\v p_2,\cdots,\v p_{k-1}}
-  T_{\v m\v p_1} T_{\v p_1\v p_2} T_{\v p_2\v p_3}
-  \cdots T_{\v p_{k-1}\v n},
+  T_{\v m\v p_{k-1}} \cdots T_{\v p_3\v p_2}
+  T_{\v p_2\v p_1} T_{\v p_1\v n},
 \end{align}
@@
 \begin{align}
-  \S_{\v m}\p{t}
+  \O\p{t}
   \to \sum_{k=0}^M \f{t^k}{k!}
-  \sum_{\v n} T^k_{\v m\v n} \S_{\v n}\p{0}.
+  \sum_{\v m, \v n} \S_{\v m} T^k_{\v m\v n} \O_{\v n}\p{0}.
   \label{eq:order_truncation}
@@
 which operators are relevant for computing the expectation value
-$\bk{\S_{\v m}\p{t}}$ to a fixed order $M$.  Note that the relation
-$\S_{\v m^*}=\S_{\v m}^\dag$, which by Hermitian conjugation of
-Eq.~\eqref{eq:EOM} also implies that
+$\bk{\O\p{t}}$ to a fixed order $M$.  Note that the relation
+$\S_{\v m}^\dag=\S_{\v m^*}$ for $\v m^*\equiv\p{m_-,m_\z,m_+}$, which
+by Hermitian conjugation of Eq.~\eqref{eq:EOM} also implies that
 $T_{\v m^*\v n^*}=T_{\v m\v n}^*$, can be exploited to cut both the
-number of initial-time expectation values $\bk{\S_{\v m}\p{0}}$ and
+number of initial-time expectation values $\bk{\S_{\v m}}_{t=0}$ and
 the number of matrix elements $T_{\v m\v n}$ that we may need to
@@
 truncated Taylor expansion in Eq.~\eqref{eq:order_truncation} only
-allows computation of the correlator $\bk{\S_{\v m}\p{t}}$ to a
+allows computation of the expectation value $\bk{\O\p{t}}$ to a
 maximal time $t_{M_{\t{max}}}$.  As we will see in the following
@@
 departure occurs because the truncated Taylor expansion neglects
-high-weight (i.e.~large-$\abs{\v n}$) operators $\S_{\v n}$ whose
-contributions to a correlator of interest $\bk{\S_{\v m}}$ eventually
+high-weight (i.e.~large-$\abs{\v m}$) operators $\S_{\v m}$ whose
+contributions to a correlator of interest $\bk{\O\p{t}}$ eventually
 become non-negligible.  For sufficiently large times $t$, these
 individual contributions may generally be large compared to the value
-of $\bk{\S_{\v m}}$, in which case their truncation will yield
+of $\bk{\O\p{t}}$, in which case their truncation will yield
 non-physical results.  This breakdown mechanism is also the reason for
@@
 rather than $\ket{\X}$: for initial states $\ket{-\Z}$, all
-initial-time correlators $\bk{\S_{\v n}\p{0}}$ vanish unless
-$n_+=n_-=0$ (see Appendix \ref{sec:initial_conditions}), so there is a
-substantially smaller number of terms with non-zero contribution to
-the correlator $\bk{\S_{\v m}}$, and therefore a substantially smaller
-number of non-zero terms that are neglected in the truncated Taylor
-expansion.
+initial-time correlators $\bk{\S_{\v m}}_{t=0}$ vanish unless
+$m_+=m_-=0$ (see Appendix \ref{sec:initial_conditions}), so there is a
+substantially smaller number of neglected terms with non-zero
+contribution to the correlator $\bk{\O\p{t}}$.
 

### 2019-03-28 soc_squeezing/collective_paper/collective_spin.tex
added some text/notes on two-time correlators commented out the last appendix on more general fixed-weight operator products

@@
 $\gamma_{\J_-}=\gamma_-$.  The commutator in Eq.~\eqref{eq:time_deriv}
-can be determined from the structure constants $f_{\v\ell\v m\v n}$ of
-the general product
-$\S_{\v\ell}\S_{\v m}=\sum_{\v n}f_{\v\ell\v m\v n}\S_{\v n}$ worked
-out in Appendices \ref{sec:identities}--\ref{sec:prod_general}, while
-the effects of decoherence from jump operators (i.e.~elements of $\J$)
-of the form $\gamma^{(j)}=\sum_n\gamma_ns_n^{(j)}$ and
+can be computed by expanding the product
+$\S_{\v\ell}\S_{\v m}=\sum_{\v n}f_{\v\ell\v m\v n}\S_{\v n}$ with
+structure constants $f_{\v\ell\v m\v n}$ that are worked out in
+Appendices \ref{sec:identities}--\ref{sec:prod_general}, and the
+effects of decoherence from jump operators (i.e.~elements of $\J$) of
+the form $\gamma^{(j)}=\sum_n\gamma_ns_n^{(j)}$ and
 $\Gamma=\sum_n\Gamma_nS_n$ are worked out in Appendices
@@
 \begin{align}
-  \f{d}{dt} \O
+  \f{d}{dt} \O\p{t}
   \to \sum_{w\p{\v m}<W} \S_{\v m}
-  \sum_{\v n} T_{\v m\v n} \O_{\v n}
+  \sum_{\v n} T_{\v m\v n} \O_{\v n}\p{t}
   \label{eq:weight_truncation}
@@
   = e^{t T} \O\p{0}
-  % = \sum_{k\ge0} \f{t^k}{k!} T^k \O\p{0}
   = \sum_{k\ge0} \f{t^k}{k!}
@@
 where the matrix elements $T^k_{\v m\v n}$ of the $k$-th time
-derivative are
+derivative operator $T^k$ are
 \begin{align}
@@
 
-\section{Two-time correlation functions}
+\section{Two-time correlation functions and non-equilibrium linear
+  response}
 \label{sec:TTC}
@@
 As a final example, we consider the calculation of two-time
-correlation functions in collective spin systems.
+correlation functions in a collective spin system.  These correlation
+functions allow us to extract the short-time linear response of
+Heisenberg operators to small perturbations of a collective spin
+system's Hamiltonian.  Consider a target Hamiltonian $H$ perturbed by
+an operator $V$, such that the net Hamiltonian is $\widetilde H=H+V$.
+We denote the generator of Heisenberg time evolution under the
+perturbed (unperturbed) Hamiltonian by $\widetilde T$ ($T$).  These
+time evolution operators are related by
+\begin{align}
+  \widetilde T = T + \hat V
+\end{align}
+where $\hat V$ is a superoperator whose action on operators $\O$ is
+defined by
+\begin{align}
+  \hat V \O \equiv i\sp{V,\O}_-.
+\end{align}
+By Taylor expanding through quadratic order in the time $t$, we can
+find that for sufficiently small times
+\begin{align}
+  e^{t\hat V} e^{t T}
+  \approx e^{t\widetilde T} - \f12 t^2 \sp{T,\hat V}_-.
+\end{align}
+To leading order in $\hat V$ and quadratic order in $t$, we therefore
+have
+\begin{align}
+  e^{t\widetilde T} - e^{tT}
+  \approx t\hat V e^{tT} + \f12 t^2 \sp{T,\hat V}_-,
+\end{align}
+where
+\begin{align}
+  \f12 t^2 \sp{T,\hat V}_- \approx \f12 t \sp{e^{tT}, \hat V}_-,
+\end{align}
+which implies
+\begin{align}
+  e^{t\widetilde T} - e^{tT} \approx \f12 t \sp{e^{tT},\hat V}_+.
+\end{align}
+Defining the perturbed and unperturbed Heisenberg operators
+$\widetilde\O\p{t}\equiv e^{t\widetilde T}\O$ and
+$\O\p{t}\equiv e^{tT}\O$, we thus find
+\begin{align}
+  \widetilde\O\p{t} - \O\p{t}
+ \approx \f12 i t \p{\sp{V,\O\p{t}}_- + \sp{V,\O}_-\p{t}}.
+\end{align}
+
+\newpage
 
@@
 
-\section{Changing operator bases, and a general product for algebras
-  of permutationally-symmetric operators}
-\label{sec:new_basis}
-
-For a collective spin system composed of many small spins, the
-operators $\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$ contain
-$\abs{\v m}$-body operators (for $\abs{\v m}\equiv\sum_\mu m_\mu$),
-but also carry lots of ``baggage'' in the form of $k$-body operators
-with $k<\abs{\v m}$.  To cut down on this overhead, we try to change
-our basis for collective spin operators from $\S_{\v m}$ to the
-purely-$\abs{\v m}$-body operators
-\begin{align}
-  \tilde\S_{\v m} \equiv \sum_{\p{\v j;\v m}} P_{\v j},
-  &&
-  P_{\v j} \equiv \prod_{\mu,a} s_\mu^{\p{j^\mu_a}},
-  \label{eq:S_P}
-\end{align}
-where $\v m \equiv \p{m_+,m_\z,m_-}\in\mathbb{N}_0^3$ with $m_\mu$ the
-number of $s_\mu$ operators in each term of $\S_{\v m}$; $\v j$ is an
-ordered list of indices for the spins addressed by $P_{\v j}$; and the
-sum over $\p{\v j;\v m}$ denotes a sum over all possible $\v j$ for
-which all indices in $\v j$ are distinct and
-$\abs{\set{j^\mu_a}} = m_\mu$, such that $j^\mu_a$ indexes the $a$-th
-spin addressed by an $s_\mu$ operator in $P_{\v j}$.
-
-In order to make use of the operators $\tilde S_{\v m}$, we will need
-to compute products of the form $\tilde S_{\v m} \tilde S_{\v n}$.  We
-therefore expand such a product into terms that have exactly $s$ spins
-addressed by both operators:
-\begin{align}
-  \tilde\S_{\v m} \tilde\S_{\v n}
-  = \sum_{s\ge0} \sum_{\p{\v j,\v k;\v m,\v n,s}} P_{\v j} P_{\v k},
-  &&
-  \sum_{\p{\v j,\v k;\v m,\v n,s}} X \equiv
-  \sum_{\substack{\p{\v j;\v m},\p{\v k;\v n} \\
-      \abs{\set{j_\alpha^\mu}\cap\set{k_\beta^\nu}}=s}} X.
-  \label{eq:SS_PP}
-\end{align}
-Collecting terms in which $r_{\mu\nu}$ of the $s_\mu$ operators in
-$P_{\v j}$ address the same spin as an $s_\nu$ operator in $P_{\v k}$,
-we have
-\begin{align}
-  \tilde\S_{\v m} \tilde\S_{\v n}
-  = \sum_{s\ge0} \sum_{\p{\v r;\v m,\v n,s}} f_{\v m\v n\v r}
-  \sum_{\p{\v J,\v K;\v m,\v n,\v r}} P_{\v J} Q_{\v K},
-  \label{eq:SS_PQ}
-\end{align}
-where the sum over $\p{\v r;\v m,\v n,s}$ denotes a sum over all
-values of $\v r$ with the restrictions
-\begin{align}
-  r_{\mu\nu} \ge 0 ~\forall~ \mu,\nu,
-  &&
-  \sum_\nu r_{\mu\nu} \le m_\mu,
-  &&
-  \sum_\mu r_{\mu\nu} \le n_\nu,
-  &&
-  \sum_{\mu,\nu} r_{\mu\nu} = s;
-  \label{eq:rest_r}
-\end{align}
-the factor $f_{\v m\v n\v r}$ counts the number of ways to pair
-operators in $P_{\v j}$ and $P_{\v k}$ according to $\v r$, or
-\begin{align}
-  f_{\v m\v n\v r}
-  &\equiv \sp{\prod_{\mu,\nu}
-    { m_\mu - \sum_{\rho<\nu} r_{\mu\rho} \choose r_{\mu\nu} }
-    { n_\mu - \sum_{\rho<\nu} r_{\rho\mu} \choose r_{\nu\mu} }}
-  \sp{\prod_{\mu,\nu} r_{\mu\nu}!} \\
-  &= \sp{\prod_\mu \f{m_\mu!}{\p{m_\mu-\sum_\rho r_{\mu\rho}}!}
-    \f{n_\mu!}{\p{n_\mu-\sum_\rho r_{\rho\mu}}!}}
-  \sp{\prod_{\mu,\nu} r_{\mu\nu}!}^{-1};
-\end{align}
-$\v J$ and $\v K$ are respectively ordered lists of indices for spins
-addressed by one and two single-spin operators in a term of the
-product $P_{\v j} P_{\v k}$ with fixed $\v r$; the sum over
-$\p{\v J,\v K;\v m,\v n,\v r}$ denotes a sum over all values of
-$\v J,\v K$ with the restrictions
-\begin{align}
-  \abs{\set{J_a^\mu}}
-  = m_\mu + n_\mu - \sum_\rho\p{r_{\mu\rho}+r_{\rho\mu}},
-  &&
-  \abs{\set{K^{\mu\nu}_a}} = r_{\mu\nu},
-  &&
-  \set{J_a^\lambda} \cap \set{K_b^{\mu\nu}} = \varnothing,
-\end{align}
-with $\varnothing$ denoting the empty set; and finally, similarly to
-$P_{\v j}$ in \eqref{eq:S_P} we define
-\begin{align}
-  Q_{\v K} \equiv \prod_{\mu,\nu,a}
-  s_\mu^{(K^{\mu\nu}_a)} s_\nu^{(K^{\mu\nu}_a)}.
-\end{align}
-As the sum in \eqref{eq:SS_PQ} is invariant under permutation of the
-indices in $\v K$, we can safely neglect keeping track of these
-indices and simply write
-\begin{align}
-  Q_{\v K}
-  = \prod_{\mu,\nu} \prod_{a=1}^{r_{\mu\nu}}
-  s_\mu^{(K^{\mu\nu}_a)} s_\nu^{(K^{\mu\nu}_a)}
-  \simeq \bigotimes_{\mu,\nu} \bigotimes_{a=1}^{r_{\mu\nu}} s_\mu s_\nu
-  = \bigotimes_{\mu,\nu} \bigotimes_{a=1}^{r_{\mu\nu}}
-  \sum_\rho \eta_{\mu\nu\rho} s_\rho,
-  \label{eq:Q_K_eta}
-\end{align}
-where $\simeq$ denotes equality up to a re-indexing of spins and
-additional tensor factors of the single-spin identity operator
-(i.e.~$\1$); we have introduced explicit dependence on
-$r_{\mu\nu}=\abs{\set{K^{\mu\nu}_a}}$ for brevity; and
-$\eta_{\mu\nu\rho}$ is a structure constant defined by
-$s_\mu s_\nu=\sum_\rho\eta_{\mu\nu\rho}s_\rho$.  Unlike the sums over
-$\mu,\nu$ in most of the work above, the sum over $\rho$ here includes
-an index for the identity operator $s_0\equiv\1$.  Distributing the
-product of sums into a sum of products gives us
-\begin{align}
-  Q_{\v K}
-  \simeq \sum_{\v\rho} \bigotimes_{\mu,\nu,a}
-  \eta_{\mu\nu\rho^{\mu\nu}_a} s_{\rho^{\mu\nu}_a}
-  = \sum_{\v\rho} \sp{\prod_{\mu,\nu,a}\eta_{\mu\nu\rho^{\mu\nu}_a}}
-  \bigotimes_{\mu,\nu,a} s_{\rho^{\mu\nu}_a},
-  \label{eq:Q_K_rho}
-\end{align}
-where implicitly $\abs{\set{\rho^{\mu\nu}_a}}=r_{\mu\nu}$, as
-$\p{\mu,\nu,a}$ essentially index a factor in \eqref{eq:Q_K_eta} and
-$\rho^{\mu\nu}_a$ indexes one of the terms in that factor.  Letting
-$\tilde\rho^{\mu\nu}_\kappa$ denote the number of elements in the
-index list
-$\p{\rho^{\mu\nu}_1,\rho^{\mu\nu}_2,\cdots,\rho^{\mu\nu}_{r_{\mu\nu}}}$
-that are equal to $\kappa$, we observe that two terms in
-\eqref{eq:Q_K_rho} with, say, $\v\rho=\v\rho_1$ and $\v\rho=\v\rho_2$
-are equal up to a permutation of indices if
-$\tilde{\v\rho}_1=\tilde{\v\rho}_2$.  The degeneracy (under
-permutation) of terms in \eqref{eq:Q_K_rho}, i.e.~the number of
-$\v\rho$ that are consistent with $\tilde{\v\rho}$, is
-\begin{align}
-  g_{\tilde{\v\rho}}
-  \equiv \prod_{\mu,\nu,\kappa}
-  { r_{\mu\nu} - \sum_{\lambda<\kappa} \tilde\rho^{\mu\nu}_\lambda
-    \choose \tilde\rho^{\mu\nu}_\kappa }
-  = \sp{\prod_{\mu,\nu} r_{\mu\nu}!}
-  \sp{\prod_{\mu,\nu,\kappa} \tilde\rho^{\mu\nu}_\kappa!}^{-1}.
-\end{align}
-Defining additionally
-\begin{align}
-  \v\eta^{\tilde{\v\rho}}
-  \equiv \prod_{\mu,\nu,\kappa}
-  \p{\eta_{\mu\nu\kappa}}^{\tilde\rho^{\mu\nu}_\kappa},
-  \label{eq:eta_rho}
-\end{align}
-we can collect like factors and group together equivalent terms in
-\eqref{eq:Q_K_rho} to write
-\begin{align}
-  Q_{\v K}
-  \simeq \sum_{\tilde{\v\rho}} g_{\tilde{\v\rho}} \v\eta^{\tilde{\v\rho}}
-  \bigotimes_{\kappa}
-  s_\kappa^{\otimes\sum_{\mu,\nu}\tilde\rho^{\mu\nu}_\kappa}
-  \simeq \sum_{\tilde{\v\rho}} g_{\tilde{\v\rho}} \v\eta^{\tilde{\v\rho}}
-  \1^{\otimes\sum_{\mu,\nu}\tilde\rho^{\mu\nu}_0}
-  \otimes \bigotimes_{\kappa\ne0}
-  s_\kappa^{\otimes\sum_{\mu,\nu}\tilde\rho^{\mu\nu}_\kappa},
-  \label{eq:Q_K_rho_simp}
-\end{align}
-where again implicitly
-$\sum_\kappa\tilde\rho^{\mu\nu}_\kappa=r_{\mu\nu}$ for consistency
-with \eqref{eq:Q_K_eta} and \eqref{eq:Q_K_rho}, and we have explicitly
-factored out the identity operators in each term of $Q_{\v K}$.
-
-We have now simplified $Q_{\v K}$ sufficiently to substitute it back
-into \eqref{eq:SS_PQ}, which (removing the tilde from $\tilde{\v\rho}$
-to simplify notation) gives us
-\begin{align}
-  \tilde\S_{\v m} \tilde\S_{\v n}
-  = \sum_{s\ge0} \sum_{\p{\v r;\v m,\v n,s}} f_{\v m\v n\v r}
-  \sum_{\p{\v\rho;\v r}} g_{\v\rho} \v\eta^{\v\rho}
-  c_{\v\ell_{\v m\v n\v r\v\rho}\v\rho}
-  \tilde\S_{\v\ell_{\v m\v n\v r\v\rho}},
-  \label{eq:SS_simp}
-\end{align}
-where the sum over $\p{\v\rho;\v r}$ denotes a sum over all values of
-$\v\rho$ and with the restrictions
-\begin{align}
-  \abs{\set{\rho^{\mu\nu}_\kappa}} = \abs{\set{s_\kappa}},
-  &&
-  \sum_{\kappa} \rho^{\mu\nu}_\kappa = r_{\mu\nu};
-\end{align}
-the combinatorial factor
-\begin{align}
-  c_{\v\ell\v\rho}
-  \equiv \f{\p{N - \abs{\v\ell}}!}
-  {\p{N - \abs{\v\ell} - \sum_{\mu,\nu}\rho^{\mu\nu}_0}!}
-\end{align}
-accounts for the number of terms that are equivalent up to a
-permutation of indices involving the identity operators in
-\eqref{eq:Q_K_rho_simp}; and the number of $s_\mu$ operators in each
-term of $\tilde\S_{\v\ell_{\v m\v n\v r\v\rho}}$ is
-\begin{align}
-  \ell_{\v m\v n\v r\v\rho;\mu}
-  = m_\mu + n_\mu - \sum_\kappa \p{r_{\mu\kappa}+r_{\kappa\mu}}
-  + \sum_{\alpha,\beta} \rho^{\alpha\beta}_\mu.
-\end{align}
-
-The sums in \eqref{eq:SS_simp} generally involve a large number of
-terms; in practice, many of these terms will be equal to zero.  We
-can, however, impose additional restrictions on these sums in such a
-way as to throw out all terms that are zero, and keep only those that
-are not.  Remembering that $s$ counts the number of spins that are
-addressed by both operators $P_{\v j}$ and $P_{\v k}$ in
-\eqref{eq:SS_PP}, the first restriction we can impose comes from
-recognizing that for given $\tilde\S_{\v m},\tilde\S_{\v n}$, the
-overlap $s$ is bounded as
-\begin{align}
-  \max\set{0,\abs{\v m}+\abs{\v n}-N}
-  \le s \le \min\set{N,\abs{\v m}+\abs{\v n}}.
-\end{align}
-In addition, the matrix of structure constants $\eta_{\mu\nu\kappa}$
-may generally contain zeros that we can preemptively avoid with
-restrictions on $\v r$ and $\v\rho$.  Recalling that $r_{\mu\nu}$
-counts the number of operators addressed by
-$s_\mu s_\nu=\sum_\kappa\eta_{\mu\nu\kappa} s_\kappa$, if
-$\eta_{\mu\nu\kappa}=0$ for all $\kappa$, i.e.~$s_\mu s_\nu=0$, then
-any term with $r_{\mu\nu}>0$ will vanish.  Our second restriction on
-the sums in \eqref{eq:SS_simp} is therefore to fix $r_{\mu\nu}=0$ for
-all $\mu,\nu$ with $s_\mu s_\nu=0$.  Finally, we can avoid the
-nullifying effect of zeros from the factor
-$\v\eta^{\v\rho}=\prod_{\mu,\nu,\kappa}
-\p{\eta_{\mu\nu\kappa}}^{\rho^{\mu\nu}_\kappa}$ by fixing
-$\rho^{\mu\nu}_\kappa=0$ for all $\mu,\nu,\kappa$ with
-$\eta_{\mu\nu\kappa}=0$.  These restrictions are sufficient to ensure
-that all individual terms in \eqref{eq:SS_simp} are nonzero, although
-they do not rule out the possibility of several terms canceling out.
-
-TODO: comment on the use of this new basis vs.~the old one.
+% \section{Changing operator bases, and a general product for algebras
+%   of permutationally-symmetric operators}
+% \label{sec:new_basis}
+
+% For a collective spin system composed of many small spins, the
+% operators $\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$ contain
+% $\abs{\v m}$-body operators (for $\abs{\v m}\equiv\sum_\mu m_\mu$),
+% but also carry lots of ``baggage'' in the form of $k$-body operators
+% with $k<\abs{\v m}$.  To cut down on this overhead, we try to change
+% our basis for collective spin operators from $\S_{\v m}$ to the
+% purely-$\abs{\v m}$-body operators
+% \begin{align}
+%   \tilde\S_{\v m} \equiv \sum_{\p{\v j;\v m}} P_{\v j},
+%   &&
+%   P_{\v j} \equiv \prod_{\mu,a} s_\mu^{\p{j^\mu_a}},
+%   \label{eq:S_P}
+% \end{align}
+% where $\v m \equiv \p{m_+,m_\z,m_-}\in\mathbb{N}_0^3$ with $m_\mu$ the
+% number of $s_\mu$ operators in each term of $\S_{\v m}$; $\v j$ is an
+% ordered list of indices for the spins addressed by $P_{\v j}$; and the
+% sum over $\p{\v j;\v m}$ denotes a sum over all possible $\v j$ for
+% which all indices in $\v j$ are distinct and
+% $\abs{\set{j^\mu_a}} = m_\mu$, such that $j^\mu_a$ indexes the $a$-th
+% spin addressed by an $s_\mu$ operator in $P_{\v j}$.
+
+% In order to make use of the operators $\tilde S_{\v m}$, we will need
+% to compute products of the form $\tilde S_{\v m} \tilde S_{\v n}$.  We
+% therefore expand such a product into terms that have exactly $s$ spins
+% addressed by both operators:
+% \begin{align}
+%   \tilde\S_{\v m} \tilde\S_{\v n}
+%   = \sum_{s\ge0} \sum_{\p{\v j,\v k;\v m,\v n,s}} P_{\v j} P_{\v k},
+%   &&
+%   \sum_{\p{\v j,\v k;\v m,\v n,s}} X \equiv
+%   \sum_{\substack{\p{\v j;\v m},\p{\v k;\v n} \\
+%       \abs{\set{j_\alpha^\mu}\cap\set{k_\beta^\nu}}=s}} X.
+%   \label{eq:SS_PP}
+% \end{align}
+% Collecting terms in which $r_{\mu\nu}$ of the $s_\mu$ operators in
+% $P_{\v j}$ address the same spin as an $s_\nu$ operator in $P_{\v k}$,
+% we have
+% \begin{align}
+%   \tilde\S_{\v m} \tilde\S_{\v n}
+%   = \sum_{s\ge0} \sum_{\p{\v r;\v m,\v n,s}} f_{\v m\v n\v r}
+%   \sum_{\p{\v J,\v K;\v m,\v n,\v r}} P_{\v J} Q_{\v K},
+%   \label{eq:SS_PQ}
+% \end{align}
+% where the sum over $\p{\v r;\v m,\v n,s}$ denotes a sum over all
+% values of $\v r$ with the restrictions
+% \begin{align}
+%   r_{\mu\nu} \ge 0 ~\forall~ \mu,\nu,
+%   &&
+%   \sum_\nu r_{\mu\nu} \le m_\mu,
+%   &&
+%   \sum_\mu r_{\mu\nu} \le n_\nu,
+%   &&
+%   \sum_{\mu,\nu} r_{\mu\nu} = s;
+%   \label{eq:rest_r}
+% \end{align}
+% the factor $f_{\v m\v n\v r}$ counts the number of ways to pair
+% operators in $P_{\v j}$ and $P_{\v k}$ according to $\v r$, or
+% \begin{align}
+%   f_{\v m\v n\v r}
+%   &\equiv \sp{\prod_{\mu,\nu}
+%     { m_\mu - \sum_{\rho<\nu} r_{\mu\rho} \choose r_{\mu\nu} }
+%     { n_\mu - \sum_{\rho<\nu} r_{\rho\mu} \choose r_{\nu\mu} }}
+%   \sp{\prod_{\mu,\nu} r_{\mu\nu}!} \\
+%   &= \sp{\prod_\mu \f{m_\mu!}{\p{m_\mu-\sum_\rho r_{\mu\rho}}!}
+%     \f{n_\mu!}{\p{n_\mu-\sum_\rho r_{\rho\mu}}!}}
+%   \sp{\prod_{\mu,\nu} r_{\mu\nu}!}^{-1};
+% \end{align}
+% $\v J$ and $\v K$ are respectively ordered lists of indices for spins
+% addressed by one and two single-spin operators in a term of the
+% product $P_{\v j} P_{\v k}$ with fixed $\v r$; the sum over
+% $\p{\v J,\v K;\v m,\v n,\v r}$ denotes a sum over all values of
+% $\v J,\v K$ with the restrictions
+% \begin{align}
+%   \abs{\set{J_a^\mu}}
+%   = m_\mu + n_\mu - \sum_\rho\p{r_{\mu\rho}+r_{\rho\mu}},
+%   &&
+%   \abs{\set{K^{\mu\nu}_a}} = r_{\mu\nu},
+%   &&
+%   \set{J_a^\lambda} \cap \set{K_b^{\mu\nu}} = \varnothing,
+% \end{align}
+% with $\varnothing$ denoting the empty set; and finally, similarly to
+% $P_{\v j}$ in \eqref{eq:S_P} we define
+% \begin{align}
+%   Q_{\v K} \equiv \prod_{\mu,\nu,a}
+%   s_\mu^{(K^{\mu\nu}_a)} s_\nu^{(K^{\mu\nu}_a)}.
+% \end{align}
+% As the sum in \eqref{eq:SS_PQ} is invariant under permutation of the
+% indices in $\v K$, we can safely neglect keeping track of these
+% indices and simply write
+% \begin{align}
+%   Q_{\v K}
+%   = \prod_{\mu,\nu} \prod_{a=1}^{r_{\mu\nu}}
+%   s_\mu^{(K^{\mu\nu}_a)} s_\nu^{(K^{\mu\nu}_a)}
+%   \simeq \bigotimes_{\mu,\nu} \bigotimes_{a=1}^{r_{\mu\nu}} s_\mu s_\nu
+%   = \bigotimes_{\mu,\nu} \bigotimes_{a=1}^{r_{\mu\nu}}
+%   \sum_\rho \eta_{\mu\nu\rho} s_\rho,
+%   \label{eq:Q_K_eta}
+% \end{align}
+% where $\simeq$ denotes equality up to a re-indexing of spins and
+% additional tensor factors of the single-spin identity operator
+% (i.e.~$\1$); we have introduced explicit dependence on
+% $r_{\mu\nu}=\abs{\set{K^{\mu\nu}_a}}$ for brevity; and
+% $\eta_{\mu\nu\rho}$ is a structure constant defined by
+% $s_\mu s_\nu=\sum_\rho\eta_{\mu\nu\rho}s_\rho$.  Unlike the sums over
+% $\mu,\nu$ in most of the work above, the sum over $\rho$ here includes
+% an index for the identity operator $s_0\equiv\1$.  Distributing the
+% product of sums into a sum of products gives us
+% \begin{align}
+%   Q_{\v K}
+%   \simeq \sum_{\v\rho} \bigotimes_{\mu,\nu,a}
+%   \eta_{\mu\nu\rho^{\mu\nu}_a} s_{\rho^{\mu\nu}_a}
+%   = \sum_{\v\rho} \sp{\prod_{\mu,\nu,a}\eta_{\mu\nu\rho^{\mu\nu}_a}}
+%   \bigotimes_{\mu,\nu,a} s_{\rho^{\mu\nu}_a},
+%   \label{eq:Q_K_rho}
+% \end{align}
+% where implicitly $\abs{\set{\rho^{\mu\nu}_a}}=r_{\mu\nu}$, as
+% $\p{\mu,\nu,a}$ essentially index a factor in \eqref{eq:Q_K_eta} and
+% $\rho^{\mu\nu}_a$ indexes one of the terms in that factor.  Letting
+% $\tilde\rho^{\mu\nu}_\kappa$ denote the number of elements in the
+% index list
+% $\p{\rho^{\mu\nu}_1,\rho^{\mu\nu}_2,\cdots,\rho^{\mu\nu}_{r_{\mu\nu}}}$
+% that are equal to $\kappa$, we observe that two terms in
+% \eqref{eq:Q_K_rho} with, say, $\v\rho=\v\rho_1$ and $\v\rho=\v\rho_2$
+% are equal up to a permutation of indices if
+% $\tilde{\v\rho}_1=\tilde{\v\rho}_2$.  The degeneracy (under
+% permutation) of terms in \eqref{eq:Q_K_rho}, i.e.~the number of
+% $\v\rho$ that are consistent with $\tilde{\v\rho}$, is
+% \begin{align}
+%   g_{\tilde{\v\rho}}
+%   \equiv \prod_{\mu,\nu,\kappa}
+%   { r_{\mu\nu} - \sum_{\lambda<\kappa} \tilde\rho^{\mu\nu}_\lambda
+%     \choose \tilde\rho^{\mu\nu}_\kappa }
+%   = \sp{\prod_{\mu,\nu} r_{\mu\nu}!}
+%   \sp{\prod_{\mu,\nu,\kappa} \tilde\rho^{\mu\nu}_\kappa!}^{-1}.
+% \end{align}
+% Defining additionally
+% \begin{align}
+%   \v\eta^{\tilde{\v\rho}}
+%   \equiv \prod_{\mu,\nu,\kappa}
+%   \p{\eta_{\mu\nu\kappa}}^{\tilde\rho^{\mu\nu}_\kappa},
+%   \label{eq:eta_rho}
+% \end{align}
+% we can collect like factors and group together equivalent terms in
+% \eqref{eq:Q_K_rho} to write
+% \begin{align}
+%   Q_{\v K}
+%   \simeq \sum_{\tilde{\v\rho}} g_{\tilde{\v\rho}} \v\eta^{\tilde{\v\rho}}
+%   \bigotimes_{\kappa}
+%   s_\kappa^{\otimes\sum_{\mu,\nu}\tilde\rho^{\mu\nu}_\kappa}
+%   \simeq \sum_{\tilde{\v\rho}} g_{\tilde{\v\rho}} \v\eta^{\tilde{\v\rho}}
+%   \1^{\otimes\sum_{\mu,\nu}\tilde\rho^{\mu\nu}_0}
+%   \otimes \bigotimes_{\kappa\ne0}
+%   s_\kappa^{\otimes\sum_{\mu,\nu}\tilde\rho^{\mu\nu}_\kappa},
+%   \label{eq:Q_K_rho_simp}
+% \end{align}
+% where again implicitly
+% $\sum_\kappa\tilde\rho^{\mu\nu}_\kappa=r_{\mu\nu}$ for consistency
+% with \eqref{eq:Q_K_eta} and \eqref{eq:Q_K_rho}, and we have explicitly
+% factored out the identity operators in each term of $Q_{\v K}$.
+
+% We have now simplified $Q_{\v K}$ sufficiently to substitute it back
+% into \eqref{eq:SS_PQ}, which (removing the tilde from $\tilde{\v\rho}$
+% to simplify notation) gives us
+% \begin{align}
+%   \tilde\S_{\v m} \tilde\S_{\v n}
+%   = \sum_{s\ge0} \sum_{\p{\v r;\v m,\v n,s}} f_{\v m\v n\v r}
+%   \sum_{\p{\v\rho;\v r}} g_{\v\rho} \v\eta^{\v\rho}
+%   c_{\v\ell_{\v m\v n\v r\v\rho}\v\rho}
+%   \tilde\S_{\v\ell_{\v m\v n\v r\v\rho}},
+%   \label{eq:SS_simp}
+% \end{align}
+% where the sum over $\p{\v\rho;\v r}$ denotes a sum over all values of
+% $\v\rho$ and with the restrictions
+% \begin{align}
+%   \abs{\set{\rho^{\mu\nu}_\kappa}} = \abs{\set{s_\kappa}},
+%   &&
+%   \sum_{\kappa} \rho^{\mu\nu}_\kappa = r_{\mu\nu};
+% \end{align}
+% the combinatorial factor
+% \begin{align}
+%   c_{\v\ell\v\rho}
+%   \equiv \f{\p{N - \abs{\v\ell}}!}
+%   {\p{N - \abs{\v\ell} - \sum_{\mu,\nu}\rho^{\mu\nu}_0}!}
+% \end{align}
+% accounts for the number of terms that are equivalent up to a
+% permutation of indices involving the identity operators in
+% \eqref{eq:Q_K_rho_simp}; and the number of $s_\mu$ operators in each
+% term of $\tilde\S_{\v\ell_{\v m\v n\v r\v\rho}}$ is
+% \begin{align}
+%   \ell_{\v m\v n\v r\v\rho;\mu}
+%   = m_\mu + n_\mu - \sum_\kappa \p{r_{\mu\kappa}+r_{\kappa\mu}}
+%   + \sum_{\alpha,\beta} \rho^{\alpha\beta}_\mu.
+% \end{align}
+
+% The sums in \eqref{eq:SS_simp} generally involve a large number of
+% terms; in practice, many of these terms will be equal to zero.  We
+% can, however, impose additional restrictions on these sums in such a
+% way as to throw out all terms that are zero, and keep only those that
+% are not.  Remembering that $s$ counts the number of spins that are
+% addressed by both operators $P_{\v j}$ and $P_{\v k}$ in
+% \eqref{eq:SS_PP}, the first restriction we can impose comes from
+% recognizing that for given $\tilde\S_{\v m},\tilde\S_{\v n}$, the
+% overlap $s$ is bounded as
+% \begin{align}
+%   \max\set{0,\abs{\v m}+\abs{\v n}-N}
+%   \le s \le \min\set{N,\abs{\v m}+\abs{\v n}}.
+% \end{align}
+% In addition, the matrix of structure constants $\eta_{\mu\nu\kappa}$
+% may generally contain zeros that we can preemptively avoid with
+% restrictions on $\v r$ and $\v\rho$.  Recalling that $r_{\mu\nu}$
+% counts the number of operators addressed by
+% $s_\mu s_\nu=\sum_\kappa\eta_{\mu\nu\kappa} s_\kappa$, if
+% $\eta_{\mu\nu\kappa}=0$ for all $\kappa$, i.e.~$s_\mu s_\nu=0$, then
+% any term with $r_{\mu\nu}>0$ will vanish.  Our second restriction on
+% the sums in \eqref{eq:SS_simp} is therefore to fix $r_{\mu\nu}=0$ for
+% all $\mu,\nu$ with $s_\mu s_\nu=0$.  Finally, we can avoid the
+% nullifying effect of zeros from the factor
+% $\v\eta^{\v\rho}=\prod_{\mu,\nu,\kappa}
+% \p{\eta_{\mu\nu\kappa}}^{\rho^{\mu\nu}_\kappa}$ by fixing
+% $\rho^{\mu\nu}_\kappa=0$ for all $\mu,\nu,\kappa$ with
+% $\eta_{\mu\nu\kappa}=0$.  These restrictions are sufficient to ensure
+% that all individual terms in \eqref{eq:SS_simp} are nonzero, although
+% they do not rule out the possibility of several terms canceling out.
+
+% TODO: comment on the use of this new basis vs.~the old one.
 

### 2019-03-28 soc_squeezing/collective_paper/collective_spin.tex
greatly simplified the derivation for short-time linear response and two-time correlators

@@
 
-\section{Two-time correlation functions and non-equilibrium linear
-  response}
-\label{sec:TTC}
+\section{Non-equilibrium linear response and two-time correlation
+  functions}
+\label{sec:NLR}
 
 As a final example, we consider the calculation of two-time
-correlation functions in a collective spin system.  These correlation
-functions allow us to extract the short-time linear response of
-Heisenberg operators to small perturbations of a collective spin
-system's Hamiltonian.  Consider a target Hamiltonian $H$ perturbed by
-an operator $V$, such that the net Hamiltonian is $\widetilde H=H+V$.
-We denote the generator of Heisenberg time evolution under the
-perturbed (unperturbed) Hamiltonian by $\widetilde T$ ($T$).  These
-time evolution operators are related by
-\begin{align}
-  \widetilde T = T + \hat V
+correlation functions, which allow us to extract the short-time linear
+response of Heisenberg operators to perturbations of a Hamiltonian.
+Consider a target Hamiltonian $H$ perturbed by an operator $V$ with
+$\norm{V}\ll\norm{H}$, such that the net Hamiltonian is
+$\widetilde H=H+V$.  We denote the generator of Heisenberg time
+evolution under the perturbed (unperturbed) Hamiltonian by
+$\widetilde T$ ($T$).  These generators are related by
+\begin{align}
+  \widetilde T = T + i\hat V
 \end{align}
@@
 \begin{align}
-  \hat V \O \equiv i\sp{V,\O}_-.
-\end{align}
-By Taylor expanding through quadratic order in the time $t$, we can
-find that for sufficiently small times
-\begin{align}
-  e^{t\hat V} e^{t T}
-  \approx e^{t\widetilde T} - \f12 t^2 \sp{T,\hat V}_-.
-\end{align}
-To leading order in $\hat V$ and quadratic order in $t$, we therefore
-have
-\begin{align}
-  e^{t\widetilde T} - e^{tT}
-  \approx t\hat V e^{tT} + \f12 t^2 \sp{T,\hat V}_-,
-\end{align}
-where
-\begin{align}
-  \f12 t^2 \sp{T,\hat V}_- \approx \f12 t \sp{e^{tT}, \hat V}_-,
-\end{align}
-which implies
-\begin{align}
-  e^{t\widetilde T} - e^{tT} \approx \f12 t \sp{e^{tT},\hat V}_+.
+  \hat V \O \equiv \sp{V,\O}_-.
+\end{align}
+Through quadratic order in the time $t$, we can say that
+\begin{align}
+  e^{t\widetilde T}
+  \approx \f12\sp{e^{tT}, e^{it\hat V}}_+,
+\end{align}
+which at first order in the perturbation $\hat V$ implies
+\begin{align}
+  e^{t\widetilde T}
+  \approx e^{tT} + \f12 i t \sp{\hat V, e^{t T}}_+.
 \end{align}
@@
 $\widetilde\O\p{t}\equiv e^{t\widetilde T}\O$ and
-$\O\p{t}\equiv e^{tT}\O$, we thus find
-\begin{align}
-  \widetilde\O\p{t} - \O\p{t}
- \approx \f12 i t \p{\sp{V,\O\p{t}}_- + \sp{V,\O}_-\p{t}}.
+$\O\p{t}\equiv e^{tT}\O$, we thus find that for sufficiently small
+times $t$ and weak perturbations $V$,
+\begin{align}
+  \widetilde\O\p{t}
+  \approx \O\p{t} + \f12 i t \p{\sp{V,\O\p{t}}_- + \sp{V,\O}_-\p{t}}.
 \end{align}

### 2019-04-01 soc_squeezing/collective_paper/collective_spin.tex
added some commentary on the figures computed expectation values for any polarized state

@@
 In this work, we present a new method to simulate short-time dynamics
-of collective spin systems with permutationally-symmetric decoherence.
-This method is based on a short-time expansion of exact solutions to
-the equations of motion for Heisenberg operators.  Evaluating this
-expansion requires using the structure constants of a collective spin
-operator algebra; the calculation of these structure constants (in
-Appendices \ref{sec:identities}--\ref{sec:prod_general}) is one of the
-main technical results of this work, which we hope will empower both
-analytical and numerical future studies of collective spin systems.
-We benchmark our method against both exact and quantum trajectory
-Monte Carlo computations of spin squeezing in accessible parameter
-regimes, highlighting both advantages and limitations of our
-short-time expansion.  Finally, we showcase an application of our
-method with computations that are inaccessible to other numerical
-methods.
+of collective spin systems with permutationally-symmetric sources of
+decoherence.  This method is based on a short-time expansion of exact
+solutions to the equations of motion for Heisenberg operators.
+Evaluating this expansion requires using the structure constants of a
+collective spin operator algebra; the calculation of these structure
+constants (in Appendices \ref{sec:identities}--\ref{sec:prod_general})
+is one of the main technical results of this work, which we hope will
+empower both analytical and numerical studies of collective spin
+systems in the future.  We benchmark our method against both exact and
+quantum trajectory Monte Carlo computations of spin squeezing in
+accessible parameter regimes, highlighting both advantages and
+limitations of our short-time expansion.  Finally, we showcase an
+application of our method with computations that are inaccessible to
+other numerical methods.
 
@@
 system of differential equations defined by Eq.~\eqref{eq:EOM}, and
-allows us to solve it using standard numerical methods.  Initial
+allows us to solve it using standard numerical methods.  Some initial
 conditions for this system of differential equations, i.e.~expectation
@@
 \item simulating a system of differential equations for a large number
-  of operators can be quite time-consuming,
+  of operators can be time-consuming,
 \item the weight measure $w$ may need to be chosen carefully, as the
@@
 \begin{align}
-  \label{eq:OAT}
-  H_{\t{OAT}} &= \chi S_\z^2, \\
-  \label{eq:TAT}
-  H_{\t{TAT}} &= \f{\chi}{3} \p{S_\z^2 - S_\y^2}, \\
-  \label{eq:TNT}
-  H_{\t{TNT}} &= \chi S_\z^2 + \Omega S_\x,
+  H_{\t{OAT}} &= \chi S_\z^2, \label{eq:OAT} \\
+  H_{\t{TAT}} &= \chi \f13 \p{S_\z^2 - S_\y^2}, \label{eq:TAT} \\
+  H_{\t{TNT}} &= \chi S_\z^2 + \Omega S_\x, \label{eq:TNT}
 \end{align}
@@
 TAT\cite{liu2011spin, huang2015twoaxis}.  For simplicity, we further
-fix $\Omega=S\chi$ (with $S\equiv N/2$ throughout this work) to the
+fix $\Omega=\chi S$ (with $S\equiv N/2$ throughout this work) to the
 critical value known to maximize the entanglement generation rate of
@@
 spontaneous decay, excitation, and dephasing of individual spins at
-rates $\gamma_-=\gamma_+=\gamma_\z=100\chi$.  Computations of this
-sort were used to put lower bounds on theoretically achievable spin
-squeezing via TAT with decoherence in
-Ref.~[\citenum{he2019engineering}], exemplifying a concrete
-application of the truncated Taylor expansion and the collective-spin
-structure constants calculated in this work.
+rates $\gamma_-=\gamma_+=\gamma_\z=100\chi$.  These results show that
+TNT is much more robust to this kind of decoherence than OAT or TAT, a
+finding that could not be deduced from the weak-decoherence
+simulations presented in Figure \ref{fig:benchmarking}.
+Strong-decoherence computations of this sort were also used to put
+lower bounds on theoretically achievable spin squeezing via TAT with
+decoherence in Ref.~[\citenum{he2019engineering}], exemplifying a
+concrete application of the truncated Taylor expansion and the
+collective-spin structure constants calculated in this work.
 
@@
 
-\section{Non-equilibrium linear response and two-time correlation
-  functions}
-\label{sec:NLR}
-
-As a final example, we consider the calculation of two-time
-correlation functions, which allow us to extract the short-time linear
-response of Heisenberg operators to perturbations of a Hamiltonian.
-Consider a target Hamiltonian $H$ perturbed by an operator $V$ with
-$\norm{V}\ll\norm{H}$, such that the net Hamiltonian is
+\section{Short-time linear response and two-time commutators}
+\label{sec:linear_response}
+
+As another example of physics that is numerically accessible via the
+truncated Taylor expansion of Heisenberg operators, we consider the
+calculation of two-time commutators, and their relation to the
+short-time linear response of Heisenberg operators to perturbations of
+a Hamiltonian.  Consider a target Hamiltonian $H$ perturbed by an
+operator $V$ with $\norm{V}\ll\norm{H}$, where $\norm{\O}$ denotes the
+operator norm of $\O$, such that the net Hamiltonian is
 $\widetilde H=H+V$.  We denote the generator of Heisenberg time
@@
   e^{t\widetilde T}
-  \approx \f12\sp{e^{tT}, e^{it\hat V}}_+,
+  \approx \f12\sp{e^{it\hat V}, e^{tT}}_+,
 \end{align}
@@
   e^{t\widetilde T}
-  \approx e^{tT} + \f12 i t \sp{\hat V, e^{t T}}_+.
-\end{align}
-Defining the perturbed and unperturbed Heisenberg operators
+  \approx e^{tT} + \f12 it \sp{\hat V, e^{t T}}_+.
+\end{align}
+Defining perturbed and unperturbed Heisenberg operators
 $\widetilde\O\p{t}\equiv e^{t\widetilde T}\O$ and
@@
 \begin{align}
-  \widetilde\O\p{t}
-  \approx \O\p{t} + \f12 i t \p{\sp{V,\O\p{t}}_- + \sp{V,\O}_-\p{t}}.
-\end{align}
+  \widetilde\O\p{t} - \O\p{t}
+  \approx \f12 i t \p{\sp{V,\O\p{t}}_- + \sp{V,\O}_-\p{t}}.
+  \label{eq:response}
+\end{align}
+Short-time linear response functions of this sort,
+i.e.~$\widetilde\O\p{t}-\O\p{t}$, can thus be used to experimentally
+probe two-time commutators $\sp{V,\O\p{t}}_-$ that contain information
+about the spreading of information across many-body degrees of freedom
+in a collective spin system.  The numerical methods presented in this
+paper can in turn benchmark such experiments, albeit at short times,
+even in the presence of strong decoherence.
 
@@
 
-% \section{Out-of-time-ordered correlators}
-% \label{sec:OTOC}
-
-% Compute some OTOCs?  There is a small chance that doing this will
-% take a prohibitively long amount of time...
-
-% Possibility: compare ``true'' OTOCs with decoherence
-% vs.~``practical'' OTOCs in experiments with decoherence and
-% time-reversal of coherent dynamics.
-
 \section{Conclusions}
+
+% extension to spin-boson and cavity systems / dicke model / chaos?
+% different initial conditions <--> pseudo-steady states in models with a large separation of scales
+% OTOCs; theoretical interest; comparing "true" OTOCs against "practical" ones / time reversal of coherent dynamics
 
@@
 \eqref{eq:push_z_mu_Ss}.  Up to Hermitian conjugation, the unique
-cases are, for $\mu,\nu\in\set{+1,-1}$,
+cases are, for $S\equiv N/2$ and $\mu,\nu\in\set{+1,-1}$,
 \begin{align}
@@
 
-Initial conditions for all collective spin operators with respect to
-(spin-polarized) Gaussian states in the Dicke manifold spanned by
-$\set{\ket{m}:S_\z\ket{m}=m\ket{m}}$ can be determined from the
-expansions
+Here compute the expectation values of all collective spin operators
+with respect to spin-polarized (Gaussian) states.  These states are
+parameterized by polar and azimuthal angles $\theta\in[0,\pi)$,
+$\phi\in[0,2\pi)$, and lie within the Dicke manifold spanned by states
+$\ket{k}\propto S_+^{S+k}\ket{\dn}^{\otimes N}$ with
+$S\equiv N/2$ and $S_\z\ket{k}=k\ket{k}$:
+\begin{align}
+  \ket{\theta,\phi}
+  \equiv \sp{\cos\p{\theta/2} e^{i\phi/2} \ket\up
+    + \sin\p{\theta/2} e^{-i\phi/2} \ket\dn}^{\otimes N}
+  = \sum_{k=-S}^S { N \choose S+k }^{1/2}
+  \cos\p{\theta/2}^{S+k} \sin\p{\theta/2}^{S-k} e^{ik\phi} \ket{k}.
+\end{align}
+In order to compute expectation values of collective spin operators we
+expand
 \begin{align}
@@
   S_\mu = \sum_{k=-S+\delta_{\mu,-1}}^{S-\delta_{\mu,1}}
-  g_\mu\p{S,k} \op{k+\mu}{k},
-\end{align}
-with $\mu\in\set{+1,-1}$ and
-\begin{align}
-  g_\mu\p{S,k} \equiv \sqrt{\p{S-\mu k}\p{S+\mu k+1}}.
-\end{align}
-Given the initial states
-\begin{align}
-  \ket{\pm\Z} \equiv \ket{\pm S},
-  &&
-  \ket{\X} \equiv 2^{-N/2} \sum_{k=-S}^S
-  { N \choose S+k }^{1/2} \ket{k},
-\end{align}
-where $S_\x\ket{\X}=S\ket{\X}$, in terms of $\nu\in\set{+1,-1}$ and
-$\bmu\equiv-\mu$ we then have
+  g_\mu\p{k} \op{k+\mu}{k}
+  = \sum_{k=-S+\delta_{\bmu,-1}}^{S-\delta_{\bmu,1}}
+  g_\bmu\p{k} \op{k}{k+\bmu},
+\end{align}
+with $\bmu\equiv-\mu\in\set{+1,-1}$ and
+\begin{align}
+  g_\mu\p{k} \equiv \sqrt{\p{S-\mu k}\p{S+\mu k+1}},
+\end{align}
+which implies
+\begin{align}
+  S_\mu^\ell \p{\mu S_\z^m} S_\bmu^n
+  &= \sum_{k=-S+\delta_{\mu,-1}\max\set{\ell,n}}
+  ^{S-\delta_{\mu,1}\max\set{\ell,n}} \p{\mu k}^m
+  \sp{\prod_{p=0}^{\ell-1} g_\mu\p{k+\mu p}}
+  \sp{\prod_{q=0}^{n-1} g_\mu\p{k+\mu q}}
+  \op{k+\mu\ell}{k+\mu n} \\
+  &= \sum_{\mu k=-\mu S-\delta_{\mu,-1}\max\set{\ell,n}}
+  ^{\mu S-\delta_{\mu,1}\max\set{\ell,n}} \p{\mu k}^m
+  \f{\p{S-\mu k}!}{\p{S+\mu k}!}
+  \sp{\f{\p{S+\mu k+\ell}!}{\p{S-\mu k-\ell}!}
+    \f{\p{S+\mu k+n}!}{\p{S-\mu k-n}!}}^{1/2}
+  \op{k+\mu\ell}{k+\mu n} \\
+  &= \sum_{k=-S}^{S-\max\set{\ell,n}} k^m
+  \f{\p{S-k}!}{\p{S+k}!}
+  \sp{\f{\p{S+k+\ell}!}{\p{S-k-\ell}!}
+    \f{\p{S+k+n}!}{\p{S-k-n}!}}^{1/2}
+  \op{\mu\p{k+\ell}}{\mu\p{k+n}}
+\end{align}
+We therefore have that
+\begin{align}
+  \bk{\theta,\phi|S_\mu^\ell \p{\mu S_\z^m} S_\bmu^n|\theta,\phi}
+  &= e^{i\phi \mu\p{\ell-n}} N! \sum_{k=-S}^{S-\max\set{\ell,n}}
+  \f{k^m \p{S-k}! f_{\mu\ell n}\p{k,\theta}}
+  {\p{S+k}! \p{S-k-\ell}! \p{S-k-n}!} \\
+  &= e^{i\phi \mu\p{\ell-n}} \p{-1}^m N! \sum_{k=0}^{N-\max\set{\ell,n}}
+  \f{\p{S-k}^m \p{N-k}! \tilde f_{\mu\ell n}\p{k,\theta}}
+  {k! \p{N-\ell-k}! \p{N-n-k}!}
+\end{align}
+for
+\begin{align}
+  f_{\mu\ell n}\p{k,\theta}
+  \equiv \cos\p{\theta/2}^{N+\mu\p{2k+\ell+n}}
+  \sin\p{\theta/2}^{N-\mu\p{2k+\ell+n}},
+\end{align}
+\begin{align}
+  \tilde f_{\mu\ell n}\p{k,\theta}
+  \equiv f_{\mu\ell n}\p{k-S,\theta}
+  = \cos\p{\theta/2}^{2N\delta_{\mu,-1}+\mu\p{2k+\ell+n}}
+  \sin\p{\theta/2}^{2N\delta_{\mu,1}-\mu\p{2k+\ell+n}}.
+\end{align}
+Defining the states
+\begin{align}
+  \ket{+\Z} \equiv \ket{0,0} = \ket\up^{\otimes N}, &&
+  \ket{-\Z} \equiv \ket{\pi,0} = \ket\dn^{\otimes N}, &&
+  \ket{\X} \equiv \ket{\pi/2,0}
+  = \p{\f{\ket\up+\ket\dn}{\sqrt2}}^{\otimes N},
+\end{align}
+some particular expectation values of interest are
 \begin{align}
@@
     \delta_{n,0} \p{-S}^m & \mu \ne \nu,
-  \end{cases}
+  \end{cases},
 \end{align}
@@
   \bk{\X|S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n|\X}
-  &=  \f1{2^N} \sum_{k=-S+\delta_{\mu,-1}\max\set{\ell,n}}
-  ^{S-\delta_{\mu,1}\max\set{\ell,n}} \p{\mu k}^m
-  \sp{{ N \choose S+k+\mu\ell } { N \choose S+k+\mu n }}^{1/2}
-  \notag \\
-  &\qquad\qquad\qquad\qquad
-  \times \prod_{p=0}^{\ell-1} g_\mu\p{S,k+\mu p}
-  \prod_{q=0}^{n-1} g_\bmu\p{S,k+\mu\sp{n-q}} \\
-  &= \f{N!}{2^N} \sum_{\mu k=-\mu S-\delta_{\mu,-1}\max\set{\ell,n}}
-  ^{\mu S-\delta_{\mu,1}\max\set{\ell,n}}
-  \f{\p{\mu k}^m \p{S-\mu k}!}
-  {\p{S+\mu k}!\p{S-\ell-\mu k}!\p{S-n-\mu k}!} \\
-  &= \f{N!}{2^N} \sum_{k=-S}^{S-\max\set{\ell,n}}
-  \f{k^m \p{S-k}!}{\p{S+k}!\p{S-\ell-k}!\p{S-n-k}!} \\
-  &= \p{-1}^m\f{N!}{2^N} \sum_{k=0}^{N-\max\set{\ell,n}}
+  = \p{-1}^m \f{N!}{2^N} \sum_{k=0}^{N-\max\set{\ell,n}}
   \f{\p{S-k}^m \p{N-k}!}{k!\p{N-\ell-k}!\p{N-n-k}!}.
-\end{align}
-Furthermore, using the fact that
-\begin{align}
-  e^{i\phi S_\z} S_\mu e^{-i\phi S_\z} = e^{i\mu\phi} S_\mu,
-\end{align}
-we can define the state
-\begin{align}
-  \ket{\phi_{\x\y}} \equiv e^{-i\phi S_\z}\ket{\X},
-  &&
-  \p{\cos\phi S_\x + \sin\phi S_\y} \ket{\phi_{\x\y}}
-  = S \ket{\phi_{\x\y}},
-\end{align}
-and compute
-\begin{align}
-  \bk{\phi_{\x\y}|S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n|\phi_{\x\y}}
-  = e^{i\mu\p{\ell-n}\phi} \bk{\X|S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n|\X}.
 \end{align}

### 2019-04-04 soc_squeezing/collective_paper/collective_spin.tex
some wording changes in the collective spin paper

@@
 allows us to solve it using standard numerical methods.  Some initial
-conditions for this system of differential equations, i.e.~expectation
-values of collective spin operators with respect to states that are
-simple to prepare experimentally, are provided in Appendix
-\ref{sec:initial_conditions}.
+conditions for this system of differential equations, namely
+expectation values of collective spin operators with respect to
+spin-polarized (Gaussian) states that are generally simple to prepare
+experimentally, are provided in Appendix \ref{sec:initial_conditions}.
 
@@
 
-Here compute the expectation values of all collective spin operators
+Here we compute the expectation values of collective spin operators
 with respect to spin-polarized (Gaussian) states.  These states are
@@
 $\phi\in[0,2\pi)$, and lie within the Dicke manifold spanned by states
-$\ket{k}\propto S_+^{S+k}\ket{\dn}^{\otimes N}$ with
-$S\equiv N/2$ and $S_\z\ket{k}=k\ket{k}$:
+$\ket{k}\propto S_+^{S+k}\ket{\dn}^{\otimes N}$ with $S\equiv N/2$ and
+$S_\z\ket{k}=k\ket{k}$:
 \begin{align}
@@
 \end{align}
-In order to compute expectation values of collective spin operators we
-expand
+We can likewise expand, within the Dicke manifold,
 \begin{align}
@@
 \end{align}
-with $\bmu\equiv-\mu\in\set{+1,-1}$ and
+where $\bmu\equiv-\mu\in\set{+1,-1}$ and
 \begin{align}
@@
     \f{\p{S+k+n}!}{\p{S-k-n}!}}^{1/2}
-  \op{\mu\p{k+\ell}}{\mu\p{k+n}}
+  \op{\mu\p{k+\ell}}{\mu\p{k+n}}.
 \end{align}
@@
   \f{\p{S-k}^m \p{N-k}! \tilde f_{\mu\ell n}\p{k,\theta}}
-  {k! \p{N-\ell-k}! \p{N-n-k}!}
-\end{align}
-for
+  {k! \p{N-k-\ell}! \p{N-k-n}!}
+\end{align}
+with
 \begin{align}
@@
   = \p{-1}^m \f{N!}{2^N} \sum_{k=0}^{N-\max\set{\ell,n}}
-  \f{\p{S-k}^m \p{N-k}!}{k!\p{N-\ell-k}!\p{N-n-k}!}.
+  \f{\p{S-k}^m \p{N-k}!}{k!\p{N-k-\ell}!\p{N-k-n}!}.
 \end{align}

### 2019-04-05 soc_squeezing/collective_paper/collective_spin.tex
changed notations for dissipation operators and fixed sign error in expectation value calculations

@@
 \begin{align}
-  T \O \equiv i \sp{H, \O}_- + \sum_\J \gamma_\J \D\p{\J; \O},
+  T \O \equiv i \sp{H, \O}_- + \sum_\J \gamma_\J \D\p{\J} \O,
   \label{eq:time_deriv}
@@
 where $\sp{X,Y}_\pm\equiv XY\pm YX$; $\J$ is a set of jump operators
-with a corresponding decoherence rate $\gamma_\J$; and
-\begin{align}
-  \D\p{\J;\O}
-  \equiv \sum_{J\in\J} \p{J^\dag \O J - \f12\sp{J^\dag J,\O}_+}
-\end{align}
-is the Heisenberg dissipator, or Lindblad superoperator, for all jump
-operators $J\in\J$.  Decoherence via uncorrelated decay of individual
-spins at a rate $\gamma_-$, for example, would be described by the set
-of jump operators $\J_-=\set{s_-^{(j)}:j=1,2,\cdots,N}$ with
+with a corresponding decoherence rate $\gamma_\J$; and is $\D\p{\J}$
+is a Heisenberg-picture dissipator, or Lindblad superoperator, defined
+by
+\begin{align}
+  \D\p{\J} \O
+  \equiv \sum_{J\in\J} \p{J^\dag \O J - \f12\sp{J^\dag J,\O}_+}.
+\end{align}
+Decoherence via uncorrelated decay of individual spins at a rate
+$\gamma_-$, for example, would be described by the set of jump
+operators $\J_-=\set{s_-^{(j)}:j=1,2,\cdots,N}$ with
 $\gamma_{\J_-}=\gamma_-$.  The commutator in Eq.~\eqref{eq:time_deriv}
@@
 \ref{sec:sandwich_single}--\ref{sec:decoherence_collective}.  We
-consider these calculations to be the main technical contributions of
-our work, with potential applications beyond the short-time simulation
-method presented here.  These ingredients are sufficient to compute
-matrix elements $T_{\v m\v n}$ of the time derivative operator $T$ in
-Eq.~\eqref{eq:time_deriv} in most cases of practical interest.
+consider these calculations to be some of the main technical
+contributions of this work, with potential applications beyond the
+short-time simulation method presented here.  These ingredients are
+sufficient to compute matrix elements $T_{\v m\v n}$ of the time
+derivative operator $T$ in Eq.~\eqref{eq:time_deriv} in most cases of
+practical interest.
 
@@
 by Hermitian conjugation of Eq.~\eqref{eq:EOM} also implies that
-$T_{\v m^*\v n^*}=T_{\v m\v n}^*$, can be exploited to cut both the
-number of initial-time expectation values $\bk{\S_{\v m}}_{t=0}$ and
-the number of matrix elements $T_{\v m\v n}$ that we may need to
-explicitly compute roughly in half.
+$T_{\v m^*\v n^*}=T_{\v m\v n}^*$, can cut both the number of
+initial-time expectation values $\bk{\S_{\v m}}_{t=0}$ and the number
+of matrix elements $T_{\v m\v n}$ that we may need to explicitly
+compute roughly in half.
 
@@
 decoherence of individual spins on collective spin operators of the
-form $S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$.  For shorthand, we define
-\begin{align}
-  \D\p{\gamma;\O} \equiv \D\p{\set{\gamma^{(j)}};\O}
+form $S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$.  For compactness, we define
+\begin{align}
+  \D\p{\gamma} \O
+  \equiv \D\p{\set{\gamma^{(j)}:j=1,2,\cdots,N}} \O
   = \sum_j\p{{\gamma^{(j)}}^\dag \O \gamma^{(j)}
@@
 \end{align}
-where $\gamma$ is an operator that acts on a single spin, and
+where $\gamma$ is an operator that acts on a single spin,
 $\gamma^{(j)}$ is an operator that acts with $\gamma$ on spin $j$ and
-the identity $\1$ on all other spins.
+trivially on all other spins, and $N$ is the total number of spins.
 
@@
 \begin{align}
-  \D\p{s_\mu; \O}
+  \D\p{s_\mu} \O
   = \sum_j\p{s_\bmu^{(j)} \O s_\mu^{(j)}
@@
 \begin{align}
-  \D\p{s_\bmu; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
+  \D\p{s_\bmu} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
   = S_\mu^\ell \p{S+\mu S_\z}\p{-1+\mu S_\z}^m S_\bmu^n
@@
 \begin{align}
-  \D\p{s_\mu; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
+  \D\p{s_\mu} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
   &= S_\mu^\ell \p{S - \ell - n - \mu S_\z} \p{1+\mu S_\z}^m S_\bmu^n
@@
 \begin{align}
-  \D\p{s_\z; \O}
+  \D\p{s_\z} \O
   = \sum_j\p{s_\z^{(j)} \O s_\z^{(j)}
@@
 \begin{align}
-  \D\p{s_\z; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
+  \D\p{s_\z} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
   = -\f12\p{\ell+n} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n
@@
 \begin{align}
-  \D\p{\gamma;\O}
+  \D\p{\gamma} \O
   = \sum_j\p{{\gamma^{(j)}}^\dag \O \gamma^{(j)}
@@
 \begin{align}
-  \D\p{\gamma;\O}
-  &= \sum_{X\in\set{\z,+,-}} \abs{\gamma_X}^2 \D\p{s_X;\O}
+  \D\p{\gamma} \O
+  &= \sum_{X\in\set{\z,+,-}} \abs{\gamma_X}^2 \D\p{s_X} \O
   + \sum_{\mu,j}
@@
 \begin{align}
-  \D\p{\gamma; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
+  \D\p{\gamma} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
   = \sum_{X\in\set{\z,+,-}} \abs{\gamma_X}^2
-  \D\p{s_X; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
+  \D\p{s_X} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
   + Q_{\ell mn;\mu}^{(\gamma)} + \sp{Q_{nm\ell;\mu}^{(\gamma)}}^\dag.
@@
 \begin{align}
-  \D\p{\Gamma;\O}
-  \equiv \D\p{\set{\Gamma};\O}
+  \D\p{\Gamma} \O
+  \equiv \D\p{\set{\Gamma}} \O
   = \Gamma^\dag \O \Gamma - \f12\sp{\Gamma^\dag \Gamma, \O}_+,
@@
 \begin{align}
-  \D\p{S_\bmu; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
-  = -S_\mu^{\ell+1} \sp{\p{1+\mu S_\z}^m - \p{\mu S_\z}^m} S_\bmu^{n+1}
-  + \f12 S_\mu^\ell \sp{\ell\p{\ell-1} + n\p{n-1}
+  \D\p{S_\bmu} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
+  &= -S_\mu^{\ell+1} \sp{\p{1+\mu S_\z}^m - \p{\mu S_\z}^m} S_\bmu^{n+1}
+  \notag \\
+  &\qquad + \f12 S_\mu^\ell \sp{\ell\p{\ell-1} + n\p{n-1}
     + 2\p{\ell+n}\mu S_\z} \p{\mu S_\z}^m S_\bmu^n,
@@
 \begin{align}
-  \D\p{S_\mu; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
-  &= S_\mu^{\ell+1} \sp{\p{2+\mu S_\z}^m-\p{1+\mu S_\z}^m}
-  S_\bmu^{n+1} \notag \\
+  \D\p{S_\mu} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
+  &= S_\mu^{\ell+1} \sp{\p{2+\mu S_\z}^m-\p{1+\mu S_\z}^m} S_\bmu^{n+1}
+  \notag \\
   &\qquad - S_\mu^\ell
@@
 \begin{align}
-  \D\p{S_\z; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
+  \D\p{S_\z} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
   = -\f12 \p{\ell-n}^2 S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n.
@@
 \begin{align}
-  \D\p{\Gamma;\O}
-  &= \sum_{X\in\set{\z,+,-}} \abs{\Gamma_X}^2 \D\p{S_X;\O}
+  \D\p{\Gamma} \O
+  &= \sum_{X\in\set{\z,+,-}} \abs{\Gamma_X}^2 \D\p{S_X} \O
   + \sum_\mu \p{\Gamma_\bmu^* \Gamma_\mu S_\mu \O S_\mu
@@
 \begin{align}
-  \D\p{\Gamma; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
+  \D\p{\Gamma} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
   = \sum_{X\in\set{\z,+,-}} \abs{\Gamma_X}^2
-  \D\p{S_X; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
+  \D\p{S_X} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
   + \tilde Q_{\ell mn;\mu}^{(\Gamma)}
@@
   \ket{\theta,\phi}
-  \equiv \sp{\cos\p{\theta/2} e^{i\phi/2} \ket\up
-    + \sin\p{\theta/2} e^{-i\phi/2} \ket\dn}^{\otimes N}
+  \equiv \sp{\cos\p{\theta/2} e^{-i\phi/2} \ket\up
+    + \sin\p{\theta/2} e^{i\phi/2} \ket\dn}^{\otimes N}
   = \sum_{k=-S}^S { N \choose S+k }^{1/2}
-  \cos\p{\theta/2}^{S+k} \sin\p{\theta/2}^{S-k} e^{ik\phi} \ket{k}.
+  \cos\p{\theta/2}^{S+k} \sin\p{\theta/2}^{S-k} e^{-ik\phi} \ket{k}.
 \end{align}

### 2019-04-11 soc_squeezing/collective_paper/collective_spin.tex
general revision of intro, theory, and squeezing sections; moved material in section IV to an appendix, wrote/motivated a new section IV on two-time correlators added two-time correlator figure

@@
 
-\title{Structure constants and short-time simulations of open
-  collective quantum spin systems}
+\title{Short-time simulations of open collective quantum spin systems}
 
@@
   ma2011quantum, he2019engineering} that allow for measurement errors
-$\Delta\phi\sim1/N^\epsilon$ with $\epsilon>1/2$.  Generating such
-states typically requires nonlinear dynamics that are realized using
-collisional, photon-mediated, or phonon-mediated interactions.
-Although a truly collective spin model requires uniform, all-to-all
-interactions, as long as measurements do not distinguish between
-constituent particles, even non-uniform systems can be effectively
-described by a uniform model with renormalized
-parameters\cite{hu2015entangled}.
+$\Delta\phi\sim1/N^\varepsilon$ with $1/2<\varepsilon\le1$, where
+$\varepsilon=1$ saturates the Heisenberg
+limit\cite{zwierz2010general}.  Generating such states typically
+requires nonlinear dynamics that are realized using collisional,
+photon-mediated, or phonon-mediated interactions.  Although a truly
+collective spin model requires uniform, all-to-all interactions, as
+long as measurements do not distinguish between constituent particles,
+even non-uniform systems can be effectively described by a uniform
+model with renormalized parameters\cite{hu2015entangled}.
 
@@
 \end{align}
-The time evolution of a Heisenberg operator $\O\p{t}$ under a
-Hamiltonian $H$ is then given by
+The time evolution of a Heisenberg operator $\O\p{t}$ is then
+determined by
 \begin{align}
@@
 for a time derivative (super-)operator $T$ with matrix elements
-$T_{\v m\v n}$, determined by
-\begin{align}
-  T \O \equiv i \sp{H, \O}_- + \sum_\J \gamma_\J \D\p{\J} \O,
+$T_{\v m\v n}$ defined by
+\begin{align}
+  T \S_{\v n} \equiv i \sp{H, \S_{\v n}}_-
+  + \sum_\J \gamma_\J \D\p{\J} \S_{\v n}
+  = \sum_{\v m} \S_{\v m} T_{\v m\v n},
   \label{eq:time_deriv}
 \end{align}
-where $\sp{X,Y}_\pm\equiv XY\pm YX$; $\J$ is a set of jump operators
-with a corresponding decoherence rate $\gamma_\J$; and is $\D\p{\J}$
-is a Heisenberg-picture dissipator, or Lindblad superoperator, defined
-by
+where $\sp{X,Y}_\pm\equiv XY\pm YX$; $H$ is the collective-spin
+Hamiltonian; $\J$ is a set of jump operators with a corresponding
+decoherence rate $\gamma_\J$; and is $\D\p{\J}$ is a
+Heisenberg-picture dissipator, or Lindblad superoperator, defined by
 \begin{align}
@@
 The time derivative operator $T$ will generally couple spin operators
-$\S_{\v n}$ (i.e.~in $\O$) to spin operators $\S_{\v m}$ with higher
-``weight'', i.e.~with $\abs{\v m}>\abs{\v n}$, where
+$\S_{\v n}$ to spin operators $\S_{\v m}$ with higher ``weight'',
+i.e.~with $\abs{\v m}>\abs{\v n}$, where
 $\abs{\v\ell}\equiv\sum_\alpha\ell_\alpha$.  The growth of operator
@@
 \end{align}
-for some weight measure $w$, e.g.~$w\p{\v m}=\abs{\v m}$, and cutoff
-$W$.  The truncation in Eq.~\eqref{eq:weight_truncation} closes the
-system of differential equations defined by Eq.~\eqref{eq:EOM}, and
-allows us to solve it using standard numerical methods.  Some initial
-conditions for this system of differential equations, namely
-expectation values of collective spin operators with respect to
-spin-polarized (Gaussian) states that are generally simple to prepare
-experimentally, are provided in Appendix \ref{sec:initial_conditions}.
+for some weight measure $w$, e.g.~$w\p{\v m}=\abs{\v m}$, and a
+high-weight cutoff $W$.  The truncation in
+Eq.~\eqref{eq:weight_truncation} closes the system of differential
+equations defined by Eq.~\eqref{eq:EOM}, and allows us to solve it
+using standard numerical methods.  Some initial conditions for this
+system of differential equations, namely expectation values of
+collective spin operators with respect to spin-polarized (Gaussian)
+states that are generally simple to prepare experimentally, are
+provided in Appendix \ref{sec:initial_conditions}.
 
@@
 \item the weight measure $w$ may need to be chosen carefully, as the
-  optimal measure is generally system-dependent,
+  optimal measure is generally system-dependent, and
 \item simulation results can only be trusted up to the time at which
@@
 which operators are relevant for computing the expectation value
-$\bk{\O\p{t}}$ to a fixed order $M$.  Note that the relation
+$\bk{\O\p{t}}$ to a fixed order $M$.  Note that using the relation
 $\S_{\v m}^\dag=\S_{\v m^*}$ for $\v m^*\equiv\p{m_-,m_\z,m_+}$, which
@@
 initial-time expectation values $\bk{\S_{\v m}}_{t=0}$ and the number
-of matrix elements $T_{\v m\v n}$ that we may need to explicitly
+qof matrix elements $T_{\v m\v n}$ that we may need to explicitly
 compute roughly in half.
@@
 \section{Spin squeezing, benchmarking, and breakdown}
+\label{sec:squeezing}
 
 To benchmark our method for computing collective spin correlators, we
-consider three collective spin models known to dynamically generate
-spin-squeezed states: the one-axis twisting (OAT), two-axis twisting
-(TAT), and twist-and-turn (TNT) models described by the collective
-spin Hamiltonians\cite{kitagawa1993squeezed, micheli2003manyparticle}
+consider three collective spin models known to generate spin-squeezed
+states: the one-axis twisting (OAT), two-axis twisting (TAT), and
+twist-and-turn (TNT) models described by the collective spin
+Hamiltonians\cite{kitagawa1993squeezed, micheli2003manyparticle}
 \begin{align}
   H_{\t{OAT}} &= \chi S_\z^2, \label{eq:OAT} \\
-  H_{\t{TAT}} &= \chi \f13 \p{S_\z^2 - S_\y^2}, \label{eq:TAT} \\
+  H_{\t{TAT}} &= \f{\chi}{3} \p{S_\z^2 - S_\y^2}, \label{eq:TAT} \\
   H_{\t{TNT}} &= \chi S_\z^2 + \Omega S_\x, \label{eq:TNT}
@@
 high-weight (i.e.~large-$\abs{\v m}$) operators $\S_{\v m}$ whose
-contributions to a correlator of interest $\bk{\O\p{t}}$ eventually
-become non-negligible.  For sufficiently large times $t$, these
-individual contributions may generally be large compared to the value
-of $\bk{\O\p{t}}$, in which case their truncation will yield
+contributions to a Heisenberg operator $\O\p{t}$ of interest
+eventually become non-negligible.  For sufficiently large times $t$,
+these individual contributions may generally be large compared to the
+value of $\bk{\O\p{t}}$, in which case their truncation will yield
 non-physical results.  This breakdown mechanism is also the reason for
@@
 simulations presented in Figure \ref{fig:benchmarking}.
-Strong-decoherence computations of this sort were also used to put
-lower bounds on theoretically achievable spin squeezing via TAT with
+Strong-decoherence computations of this sort were used to put lower
+bounds on theoretically achievable spin squeezing via TAT with
 decoherence in Ref.~[\citenum{he2019engineering}], exemplifying a
@@
 
-\section{Short-time linear response and two-time commutators}
-\label{sec:linear_response}
-
-As another example of physics that is numerically accessible via the
-truncated Taylor expansion of Heisenberg operators, we consider the
-calculation of two-time commutators, and their relation to the
-short-time linear response of Heisenberg operators to perturbations of
-a Hamiltonian.  Consider a target Hamiltonian $H$ perturbed by an
-operator $V$ with $\norm{V}\ll\norm{H}$, where $\norm{\O}$ denotes the
-operator norm of $\O$, such that the net Hamiltonian is
-$\widetilde H=H+V$.  We denote the generator of Heisenberg time
-evolution under the perturbed (unperturbed) Hamiltonian by
-$\widetilde T$ ($T$).  These generators are related by
-\begin{align}
-  \widetilde T = T + i\hat V
-\end{align}
-where $\hat V$ is a superoperator whose action on operators $\O$ is
-defined by
-\begin{align}
-  \hat V \O \equiv \sp{V,\O}_-.
-\end{align}
-Through quadratic order in the time $t$, we can say that
-\begin{align}
-  e^{t\widetilde T}
-  \approx \f12\sp{e^{it\hat V}, e^{tT}}_+,
-\end{align}
-which at first order in the perturbation $\hat V$ implies
-\begin{align}
-  e^{t\widetilde T}
-  \approx e^{tT} + \f12 it \sp{\hat V, e^{t T}}_+.
-\end{align}
-Defining perturbed and unperturbed Heisenberg operators
-$\widetilde\O\p{t}\equiv e^{t\widetilde T}\O$ and
-$\O\p{t}\equiv e^{tT}\O$, we thus find that for sufficiently small
-times $t$ and weak perturbations $V$,
-\begin{align}
-  \widetilde\O\p{t} - \O\p{t}
-  \approx \f12 i t \p{\sp{V,\O\p{t}}_- + \sp{V,\O}_-\p{t}}.
-  \label{eq:response}
-\end{align}
-Short-time linear response functions of this sort,
-i.e.~$\widetilde\O\p{t}-\O\p{t}$, can thus be used to experimentally
-probe two-time commutators $\sp{V,\O\p{t}}_-$ that contain information
-about the spreading of information across many-body degrees of freedom
-in a collective spin system.  The numerical methods presented in this
-paper can in turn benchmark such experiments, albeit at short times,
-even in the presence of strong decoherence.
-
-\newpage
+\section{Two-time correlation functions}
+\label{sec:two_time}
+
+As a final example of collective-spin physics that is numerically
+accessible via the truncated Taylor expansion of Heisenberg operators,
+we consider the calculation of collective-spin two-time correlation
+functions.  In particular, we consider the effect of decoherence on
+the short-time behavior of the two-time connected correlator
+\begin{align}
+  C\p{t}
+  \equiv \f1S\sp{\bk{S_+\p{t}S_-\p{0}}-\bk{S_+\p{t}}\bk{S_-\p{0}}}
+  \label{eq:two_time}
+\end{align}
+in the context of the squeezing models in Section \ref{sec:squeezing}.
+In an equilibrium setting, correlation functions of this sort contain
+information about the linear response of Heisenberg operators to
+perturbations of a system[CITE]\cite{CITE}; in a non-equilibrium
+setting, they contribute to short-time linear response (see Appendix
+\ref{sec:linear_response}).  Similar correlators have made appearances
+as order parameters that diagnose time-crystalline phases of
+matter\cite{tucker2018shattered}.  Related multi-time correlators,
+known as out-of-time-ordered correlators, are commonly examined for
+signatures of quantum chaos and information
+scrambling[CITE]\cite{CITE}.
+
+\begin{figure*}
+  \centering
+  \includegraphics{Sp_Sm.pdf}
+  \caption{Real (solid lines) and imaginary (dashed lines) parts of
+    the two-time connected correlator $C\p{t}$ defined in
+    Eq.~\eqref{eq:two_time} for $N=10^4$ spins initially in the
+    polarized state $\ket{\X}\propto\p{\ket\up+\ket\dn}^{\otimes N}$
+    evolving under the squeezing Hamiltonians in
+    Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT}.  Results are shown for time
+    evolution without decoherence (leftmost panel), as well as with
+    decoherence at rates $\gamma_n$ or $\Gamma_n$, respectively
+    corresponding to single-spin or collective jump operators
+    $s_n^{(j)}$ or $S_n$.  All axes in the sub-figures are shared.  At
+    short times, the correlations captured by $C\p{t}$ are suppressed
+    by single-spin decay ($\gamma_-$), collective decay ($\Gamma_-$),
+    and single-spin dephasing ($\gamma_\z$); these correlations are
+    relatively insensitive to collective dephasing ($\Gamma_\z$), and
+    can change sign or grow under single-spin ($\gamma_+$) or
+    collective ($\Gamma_+$) excitation.}
+  \label{fig:two_time}
+\end{figure*}
+
+Figure \ref{fig:two_time} shows the behavior of $C\p{t}$ for $N=10^4$
+spins, initially in the state
+$\ket\X\propto\p{\ket\up+\ket\dn}^{\otimes N}$, evolving under the
+squeezing Hamiltonians in Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT} both
+with and without various sources of decoherence.  The two-time
+correlator $C\p{t}$ degrades at short times under certain types of
+decoherence, but seems relatively stable to collective dephasing (with
+jump operator $S_\z$), and can qualitatively change behavior under in
+the presence of single-spin or collective excitation (with jump
+operators $s_+^{(j)}$ or $S_+$).  While we do not have an operational
+understanding of the correlations captured by $C\p{t}$ in the context
+of spin squeezing, Figure \ref{fig:two_time} serves as an example for
+the type of quantity and behavior that is accessible at short times
+with the truncated Taylor expansion.  This example is straightforward
+to extend to multi-time and out-of-time-ordered correlators, as well
+as equilibrium settings.
 
@@
 % extension to spin-boson and cavity systems / dicke model / chaos?
-% different initial conditions <--> pseudo-steady states in models with a large separation of scales
-% OTOCs; theoretical interest; comparing "true" OTOCs against "practical" ones / time reversal of coherent dynamics
-
-Mention/discuss extension of this work to spin-boson systems.
+% different initial conditions <--> steady states or pseudo-steady states in models with a large separation of scales
+% OTOCs; theoretical interest, scrambling; benchmarking "practical" OTOCs (with time-reversal of only coherent dynamics) against "true" OTOCs
 
@@
 
-\section{Exact results for the one-axis twisting model}
+\section{Analytical results for the one-axis twisting model}
 \label{sec:OAT}
@@
 Ref.~[\citenum{foss-feig2013nonequilibrium}] for the initial state
-$\ket{\X}$ satisfying $S_\x\ket{\X}=S\ket{\X}$ evolving under
-$H_{\t{OAT}}$, finding
+$\ket\X\propto\p{\ket\up+\ket\dn}^{\otimes N}$ with
+$S_\x\ket\X=S\ket\X$ evolving under $H_{\t{OAT}}$, finding
 \begin{align}
@@
   \Phi\p{X,t}
-  &\equiv e^{-\lambda t} \sp{\cos\p{t\sqrt{s_X^2-r}}
-    + \lambda t~\t{sinc}\p{t\sqrt{s_X^2-r}}},
+  &\equiv e^{-\lambda t} \sp{\cos\p{t\sqrt{q_X^2-r}}
+    + \lambda t~\t{sinc}\p{t\sqrt{q_X^2-r}}},
   \\
   \Psi\p{X,t}
-  &\equiv e^{-\lambda t} \p{is_X-\gamma}t~
-  \t{sinc}\p{t\sqrt{s_X^2-r}},
+  &\equiv e^{-\lambda t} \p{iq_X-\gamma}t~
+  \t{sinc}\p{t\sqrt{q_X^2-r}},
 \end{align}
@@
   &&
+  \Gamma \equiv \lambda + \f12\gamma_\z,
+  &&
   r \equiv \gamma_+ \gamma_-,
   &&
-  \Gamma \equiv \f12\gamma_\z + \lambda,
-  &&
-  s_X \equiv X + i\gamma,
-\end{align}
-In order to compute spin squeezing as measured by $\xi^2$ in
-\eqref{eq:squeezing}, we additionally need analytical expressions for
-$\bk{S_\z}$ and $\bk{S_\z^2}$.  As these operators commute with both
-the OAT Hamiltonian and the single-spin operators $\sigma_\z^{(j)}$,
-their evolution is governed entirely by decay-type decoherence (see
-Appendix \ref{sec:decay_single}), which means
+  q_X \equiv X + i\gamma,
+\end{align}
+In order to compute spin squeezing as measured by the Ramsey squeezing
+parameter $\xi^2$ defined in \eqref{eq:squeezing}, we additionally
+need analytical expressions for $\bk{S_\z}$ and $\bk{S_\z^2}$.  As
+these operators commute with both the OAT Hamiltonian and the
+single-spin operators $\sigma_\z^{(j)}$, their evolution is governed
+entirely by decay-type decoherence (see Appendix
+\ref{sec:decay_single}), which means
 \begin{align}
@@
 parameter $\xi^2$ in \eqref{eq:squeezing} at any time throughout
-evolution of the initial state $\ket{\X}$ under $H_{\t{OAT}}$, under
+evolution of the initial state $\ket\X$ under $H_{\t{OAT}}$, under
 appropriate assumptions about the relevant sources of decoherence.
@@
 \end{align}
-We therefore have that
+This expansion allows us to compute the expectation value
 \begin{align}
@@
 \end{align}
-with
+where
 \begin{align}
@@
   \ket{-\Z} \equiv \ket{\pi,0} = \ket\dn^{\otimes N}, &&
-  \ket{\X} \equiv \ket{\pi/2,0}
+  \ket\X \equiv \ket{\pi/2,0}
   = \p{\f{\ket\up+\ket\dn}{\sqrt2}}^{\otimes N},
@@
 \end{align}
+
+\section{Short-time linear response and two-time correlators}
+\label{sec:linear_response}
+
+Here, we discuss the appearance of two-time correlation functions in
+the short-time linear response of Heisenberg operators to
+perturbations of a Hamiltonian.  Consider an initial Hamiltonian $H$
+perturbed by an operator $V$ with $\norm{V}\ll\norm{H}$, where
+$\norm{\O}$ denotes the operator norm of $\O$, such that the net
+Hamiltonian is $\widetilde H=H+V$.  We denote the generator of
+Heisenberg time evolution under the perturbed (unperturbed)
+Hamiltonian by $\widetilde T$ ($T$).  These generators are related by
+\begin{align}
+  \widetilde T = T + i\hat V
+\end{align}
+where $\hat V$ is a superoperator whose action on operators $\O$ is
+defined by
+\begin{align}
+  \hat V \O \equiv \sp{V,\O}_-.
+\end{align}
+Through quadratic order in the time $t$ and linear order in the
+perturbation $\hat V$, we can say that
+\begin{align}
+  e^{t\widetilde T}
+  \approx \f12\sp{e^{tT}, e^{it\hat V}}_+
+  \approx e^{tT} + \f12 it \sp{e^{t T}, \hat V}_+.
+\end{align}
+Defining perturbed and unperturbed Heisenberg operators
+$\widetilde\O\p{t}\equiv e^{t\widetilde T}\O$ and
+$\O\p{t}\equiv e^{tT}\O$, we thus find that for sufficiently small
+times $t$ and weak perturbations $V$,
+\begin{align}
+  \widetilde\O\p{t} - \O\p{t}
+  = \p{e^{t\widetilde T} - e^{tT}} \O
+  \approx \f12 i t \p{\sp{V,\O}_-\p{t} + \sp{V,\O\p{t}}_-}.
+  \label{eq:response}
+\end{align}
+Two-time operators $V\O\p{t}$ and $\O\p{t} V$, in addition to the
+Heisenberg operators $\p{V\O}\p{t}$ and $\p{\O V}\p{t}$, thus
+determine the short-time linear response of operators $\O\p{t}$ to
+perturbations $V$ of a Hamiltonian.
 

### 2019-04-11 soc_squeezing/collective_paper/collective_spin.tex
renamed the method to 'truncated short-time (TST) expansion' added conclusion and acknowledgments

@@
 \end{align}
-Unlike the truncation in Eq.~\eqref{eq:weight_truncation}, the nonzero
-matrix elements $T^k_{\v m\v n}$ for $k=0,1,\cdots,M$ now tell us
-which operators are relevant for computing the expectation value
-$\bk{\O\p{t}}$ to a fixed order $M$.  Note that using the relation
-$\S_{\v m}^\dag=\S_{\v m^*}$ for $\v m^*\equiv\p{m_-,m_\z,m_+}$, which
-by Hermitian conjugation of Eq.~\eqref{eq:EOM} also implies that
+We refer Eq.~\eqref{eq:order_truncation} as the truncated short-time
+(TST) expansion of Heisenberg operators.  Unlike the truncation in
+Eq.~\eqref{eq:weight_truncation}, the nonzero matrix elements
+$T^k_{\v m\v n}$ for $k=0,1,\cdots,M$ in
+Eq.~\eqref{eq:order_truncation} tell us which operators are relevant
+for computing the expectation value $\bk{\O\p{t}}$ to a fixed order
+$M$.  Note that using the relation $\S_{\v m}^\dag=\S_{\v m^*}$ for
+$\v m^*\equiv\p{m_-,m_\z,m_+}$, which by Hermitian conjugation of
+Eq.~\eqref{eq:EOM} also implies that
 $T_{\v m^*\v n^*}=T_{\v m\v n}^*$, can cut both the number of
@@
 In principle, the factorial suppression of terms at higher orders of
-the Taylor expansion in Eq.~\eqref{eq:time_series} implies for any
-given time $t$, there exists a truncation order $M_t$ for which the
-truncation error in Eq.~\eqref{eq:order_truncation} is negligibly
-small.  In practice, only a maximal truncation order $M_{\t{max}}$ is
-accessible with limited computational resources, such that the
-truncated Taylor expansion in Eq.~\eqref{eq:order_truncation} only
-allows computation of the expectation value $\bk{\O\p{t}}$ to a
-maximal time $t_{M_{\t{max}}}$.  As we will see in the following
-section, $M_{\t{max}}=35$ will suffice for the computation of
-collective spin correlators up to times that are relevant e.g.~for
-spin squeezing protocols.
+the expansion in Eq.~\eqref{eq:time_series} implies for any given time
+$t$, there exists a truncation order $M_t$ for which the truncation
+error in Eq.~\eqref{eq:order_truncation} is negligibly small.  In
+practice, only a maximal truncation order $M_{\t{max}}$ is accessible
+with limited computational resources, such that the TST expansion in
+Eq.~\eqref{eq:order_truncation} only allows computation of the
+expectation value $\bk{\O\p{t}}$ to a maximal time $t_{M_{\t{max}}}$.
+As we will see in the following section, $M_{\t{max}}=35$ will suffice
+for the computation of collective spin correlators up to times that
+are relevant e.g.~for spin squeezing protocols.
 
@@
 these exact and quantum trajectory simulations will be used to
-benchmark the truncated Taylor expansion in
-Eq.~\eqref{eq:order_truncation}.
+benchmark the TST expansion in Eq.~\eqref{eq:order_truncation}.
 
@@
     using exact methods (solid lines), quantum trajectory simulations
-    (dots), and the truncated Taylor expansion in
-    Eq.~\eqref{eq:order_truncation} with $M=35$ (darker dashed lines).
-    Solid circles mark the times at which the truncated Taylor
-    expansion gives an unphysical result with $\xi^2<0$.}
+    (dots), and the TST expansion in Eq.~\eqref{eq:order_truncation}
+    with $M=35$ (darker dashed lines).  Solid circles mark the times
+    at which the TST expansion gives an unphysical result with
+    $\xi^2<0$.}
   \label{fig:benchmarking}
@@
 Hamiltonians in Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT}, as computed via
-both benchmarking simulations and the truncated Taylor expansion in
+both benchmarking simulations and the TST expansion in
 Eq.~\eqref{eq:order_truncation} with $M=35$.  Squeezing is shown for
@@
 the only effect of this rotation on the results presented in Figure
-\ref{fig:benchmarking} is to prolong the time for which the truncated
-Taylor expansion agrees with the benchmarking simulations.  The reason
-for different results in rotated basis has to do with the breakdown of
-the truncated Taylor expansion, which we discuss further below.
-
-The main lesson from Figure \ref{fig:benchmarking} is that the
-truncated Taylor expansion in Eq.~\eqref{eq:order_truncation} gives
-essentially exact results right up until a sudden and drastic
-departure that is simple to diagnose by inspection.  This drastic
-departure occurs because the truncated Taylor expansion neglects
+\ref{fig:benchmarking} is to prolong the time for which the TST
+expansion agrees with the benchmarking simulations.  The reason for
+different results in rotated basis has to do with the breakdown of the
+TST expansion, which we discuss further below.
+
+The main lesson from Figure \ref{fig:benchmarking} is that the TST
+expansion gives essentially exact results right up until a sudden and
+drastic departure that is simple to diagnose by inspection.  This
+drastic departure occurs because the TST expansion neglects
 high-weight (i.e.~large-$\abs{\v m}$) operators $\S_{\v m}$ whose
@@
 non-physical results.  This breakdown mechanism is also the reason for
-prolonged agreement between the truncated Taylor expansion and
-benchmarking methods when squeezing the initial state $\ket{-\Z}$
-rather than $\ket{\X}$: for initial states $\ket{-\Z}$, all
-initial-time correlators $\bk{\S_{\v m}}_{t=0}$ vanish unless
-$m_+=m_-=0$ (see Appendix \ref{sec:initial_conditions}), so there is a
-substantially smaller number of neglected terms with non-zero
-contribution to the correlator $\bk{\O\p{t}}$.
-
-Although the truncated Taylor expansion breaks down at short times, it
-has two key advantages over other methods to compute collective spin
-correlators.  First, computing spin correlators with the truncated
-Taylor expansion is generally much faster and requires much fewer
-computing resources than the alternatives.  The quantum trajectory
-Monte Carlo simulations performed for Figure
-\ref{fig:squeezing_incoherent}, for example, take $\sim10^4$ CPU hours
-to compute on standard modern computing hardware; the bulk of this
-time is spent performing sparse matrix multiplication, leaving little
-room to further optimize run time.  Parallelization can reduce the
-actual run time of these quantum trajectory simulations to $\sim10$
-hours, but at the cost of greatly increasing computing resource
-requirements.  The truncated Taylor expansion results in Figure
-\ref{fig:squeezing_incoherent}, meanwhile, take $\sim10$ seconds to
-compute with a single CPU on similar computing hardware.
-
-Second, the truncated Taylor expansion enables computing spin
-correlators in strong-decoherence regimes entirely inaccessible to
-other methods.  As an example, Figure \ref{fig:decoherence_strong}
-shows squeezing of $N=10^4$ spins initially in $\ket{\X}$, undergoing
-spontaneous decay, excitation, and dephasing of individual spins at
-rates $\gamma_-=\gamma_+=\gamma_\z=100\chi$.  These results show that
-TNT is much more robust to this kind of decoherence than OAT or TAT, a
+prolonged agreement between the TST expansion and benchmarking methods
+when squeezing the initial state $\ket{-\Z}$ rather than $\ket{\X}$:
+for initial states $\ket{-\Z}$, all initial-time correlators
+$\bk{\S_{\v m}}_{t=0}$ vanish unless $m_+=m_-=0$ (see Appendix
+\ref{sec:initial_conditions}), so there is a substantially smaller
+number of neglected terms with non-zero contribution to the correlator
+$\bk{\O\p{t}}$.
+
+Although the TST expansion breaks down at short times, it has two key
+advantages over other methods to compute collective spin correlators.
+First, computing spin correlators with the TST expansion is generally
+much faster and requires much fewer computing resources than the
+alternatives.  The quantum trajectory Monte Carlo simulations
+performed for Figure \ref{fig:squeezing_incoherent}, for example, take
+$\sim10^4$ CPU hours to compute on standard modern computing hardware;
+the bulk of this time is spent performing sparse matrix
+multiplication, leaving little room to further optimize run time.
+Parallelization can reduce the actual run time of these quantum
+trajectory simulations to $\sim10$ hours, but at the cost of greatly
+increasing computing resource requirements.  The TST expansion results
+in Figure \ref{fig:squeezing_incoherent}, meanwhile, take $\sim10$
+seconds to compute with a single CPU on similar computing hardware.
+
+Second, the TST expansion enables computing spin correlators in
+strong-decoherence regimes entirely inaccessible to other methods.  As
+an example, Figure \ref{fig:decoherence_strong} shows squeezing of
+$N=10^4$ spins initially in $\ket{\X}$, undergoing spontaneous decay,
+excitation, and dephasing of individual spins at rates
+$\gamma_-=\gamma_+=\gamma_\z=100\chi$.  These results show that TNT is
+much more robust to this kind of decoherence than OAT or TAT, a
 finding that could not be deduced from the weak-decoherence
@@
 decoherence in Ref.~[\citenum{he2019engineering}], exemplifying a
-concrete application of the truncated Taylor expansion and the
-collective-spin structure constants calculated in this work.
+concrete application of the TST expansion and the collective-spin
+structure constants calculated in this work.
 
@@
     dephasing of individual spins at rates
-    $\gamma_-=\gamma_+=\gamma_\z=100\chi$, computed using the
-    truncated Taylor expansion in Eq.~\eqref{eq:order_truncation} with
-    $M=35$.  Solid circles mark the times at which the truncated
-    Taylor expansion gives an unphysical result with $\xi^2<0$.}
+    $\gamma_-=\gamma_+=\gamma_\z=100\chi$, computed using the TST
+    expansion in Eq.~\eqref{eq:order_truncation} with $M=35$.  Solid
+    circles mark the times at which the TST expansion gives an
+    unphysical result with $\xi^2<0$.}
   \label{fig:decoherence_strong}
@@
 As a final example of collective-spin physics that is numerically
-accessible via the truncated Taylor expansion of Heisenberg operators,
-we consider the calculation of collective-spin two-time correlation
-functions.  In particular, we consider the effect of decoherence on
-the short-time behavior of the two-time connected correlator
+accessible via the TST expansion of Heisenberg operators, we consider
+the calculation of collective-spin two-time correlation functions.  In
+particular, we consider the effect of decoherence on the short-time
+behavior of the two-time connected correlator
 \begin{align}
@@
 decoherence, but seems relatively stable to collective dephasing (with
-jump operator $S_\z$), and can qualitatively change behavior under in
-the presence of single-spin or collective excitation (with jump
-operators $s_+^{(j)}$ or $S_+$).  While we do not have an operational
-understanding of the correlations captured by $C\p{t}$ in the context
-of spin squeezing, Figure \ref{fig:two_time} serves as an example for
-the type of quantity and behavior that is accessible at short times
-with the truncated Taylor expansion.  This example is straightforward
-to extend to multi-time and out-of-time-ordered correlators, as well
-as equilibrium settings.
+jump operator $S_\z$), and can qualitatively change behavior or grow
+in the presence of single-spin or collective excitation (with jump
+operators $s_+^{(j)}$ or $S_+$).  Figure \ref{fig:two_time} serves as
+an example for the type of quantity and behavior that is accessible at
+short times with the TST expansion.  This example is straightforward
+to extend to multi-time and out-of-time-ordered correlators (OTOCs),
+as well as equilibrium settings.
 
@@
 
-% extension to spin-boson and cavity systems / dicke model / chaos?
-% different initial conditions <--> steady states or pseudo-steady states in models with a large separation of scales
-% OTOCs; theoretical interest, scrambling; benchmarking "practical" OTOCs (with time-reversal of only coherent dynamics) against "true" OTOCs
+We have presented a new method for computing correlators in collective
+spin systems at short times.  This method is based on truncating a
+short-time expansion of Heisenberg operators, and can access
+correlators on time scales that are relevant to metrological
+applications such as spin squeezing.  In order to evaluate the
+truncated short-time (TST) expansion of Heisenberg operators, we have
+computed the structure constants of a collective spin operator
+algebra, which we hope will empower future analytical and numerical
+studies of collective spin systems.  Even though we considered only
+non-equilibrium spin-squeezing processes in this work, our method can
+be applied directly in an equilibrium setting, and is straightforward
+to generalize to systems such as trapped ions and optical cavities
+with collective spin-boson interactions.  In such contexts, our method
+may be used to benchmark the short-time effects of decoherence, or
+compute quantities such as out-of-time-ordered correlators (OTOCs)
+that diagnose the onset of quantum chaos and information scrambling.
+
+\section{Acknowledgments}
+
+We acknowledge helpful discussions with Robert Lewis-Swan and Kris
+Tucker.  This work is supported by the Air Force Office of Scientific
+Research (AFOSR) grant FA9550-18-1-0319; the AFOSR Multidisciplinary
+University Research Initiative (MURI) grant; the Defense Advanced
+Research Projects Agency (DARPA) and Army Research Office (ARO) grant
+W911NF-16-1-0576; the National Science Foundation (NSF) grant
+PHY-1820885; JILA-NSF grant PFC-173400; and the National Institute of
+Standards and Technology (NIST).
 

### 2019-04-11 soc_squeezing/collective_paper/collective_spin.tex
minor wording changes

@@
 as order parameters that diagnose time-crystalline phases of
-matter\cite{tucker2018shattered}.  Related multi-time correlators,
+matter\cite{tucker2018shattered}, and related multi-time correlators,
 known as out-of-time-ordered correlators, are commonly examined for
@@
     relatively insensitive to collective dephasing ($\Gamma_\z$), and
-    can change sign or grow under single-spin ($\gamma_+$) or
-    collective ($\Gamma_+$) excitation.}
+    can change sign or grow from spontaneous single-spin ($\gamma_+$)
+    or collective ($\Gamma_+$) excitation.}
   \label{fig:two_time}
@@
 correlator $C\p{t}$ degrades at short times under certain types of
-decoherence, but seems relatively stable to collective dephasing (with
-jump operator $S_\z$), and can qualitatively change behavior or grow
-in the presence of single-spin or collective excitation (with jump
-operators $s_+^{(j)}$ or $S_+$).  Figure \ref{fig:two_time} serves as
-an example for the type of quantity and behavior that is accessible at
-short times with the TST expansion.  This example is straightforward
-to extend to multi-time and out-of-time-ordered correlators (OTOCs),
-as well as equilibrium settings.
+decoherence, but seems relatively insensitive to collective dephasing
+(with jump operator $S_\z$), and can qualitatively change behavior or
+grow in the presence of single-spin or collective excitation (with
+jump operators $s_+^{(j)}$ or $S_+$).  Figure \ref{fig:two_time}
+serves as an example for the type of quantity and behavior that is
+accessible at short times with the TST expansion.  This example is
+straightforward to extend to multi-time and out-of-time-ordered
+correlators (OTOCs), as well as equilibrium settings.
 

### 2019-04-12 soc_squeezing/collective_paper/collective_spin.tex
added abstract, reference, and changed documentclass to revtex4-2

@@
-\documentclass[pra,reprint,longbibliography]{revtex4-1}
+\documentclass[pra,twocolumn,longbibliography]{revtex4-2}
 
@@
   linkcolor=blue,
-  filecolor=magenta,
   urlcolor=cyan,
@@
 
-% for strikeout text
-% normalem included to prevent underlining titles in the bibliography
-\usepackage[normalem]{ulem}
-
-% for leaving notes in the text
-\newcommand{\note}[1]{\textcolor{red}{#1}}
-
 \begin{document}
@@
 \begin{abstract}
-  Abstract goes here.
+  We present a new method to compute short-time expectation values in
+  collective spin systems with permutationally-symmetric decoherence.
+  Our method is based on a Taylor expansion of a formal solution to
+  the equations of motion for Heisenberg operators.  This expansion
+  can be truncated at finite order to obtain nearly exact results at
+  short times that are relevant for metrological applications such as
+  spin squeezing.  In order to evaluate the expansion for Heisenberg
+  operators, we compute the structure constants of a collective SU(2)
+  operator algebra.  We demonstrate the utility of our method by
+  computing spin squeezing and two-time correlation functions in
+  strong-decoherence regimes that are inaccessible to existing
+  numerical methods.  Our method is straightforward to generalize to
+  the case of a collective spin coupled to bosonic modes, as can occur
+  in trapped ion and optical cavity experiments, and may be used to
+  compute quantities such as out-of-time-ordered correlators that
+  contain short-time signatures of quantum chaos and information
+  scrambling.
 \end{abstract}
@@
   \centering
-  \subfloat[Squeezing under coherent dynamics
+  \subfloat[Squeezing with coherent dynamics
   \label{fig:squeezing_coherent}]
@@
 expansion gives essentially exact results right up until a sudden and
-drastic departure that is simple to diagnose by inspection.  This
-drastic departure occurs because the TST expansion neglects
-high-weight (i.e.~large-$\abs{\v m}$) operators $\S_{\v m}$ whose
-contributions to a Heisenberg operator $\O\p{t}$ of interest
-eventually become non-negligible.  For sufficiently large times $t$,
-these individual contributions may generally be large compared to the
-value of $\bk{\O\p{t}}$, in which case their truncation will yield
-non-physical results.  This breakdown mechanism is also the reason for
-prolonged agreement between the TST expansion and benchmarking methods
-when squeezing the initial state $\ket{-\Z}$ rather than $\ket{\X}$:
-for initial states $\ket{-\Z}$, all initial-time correlators
+drastic departure that can be diagnosed by inspection.  This drastic
+departure occurs because the TST expansion neglects high-weight
+(i.e.~large-$\abs{\v m}$) operators $\S_{\v m}$ whose contributions to
+a Heisenberg operator $\O\p{t}$ of interest eventually become
+non-negligible.  For sufficiently large times $t$, these individual
+contributions may generally be large compared to the value of
+$\bk{\O\p{t}}$, in which case their truncation will yield non-physical
+results.  This breakdown mechanism is also the reason for prolonged
+agreement between the TST expansion and benchmarking methods when
+squeezing the initial state $\ket{-\Z}$ rather than $\ket{\X}$: for
+initial states $\ket{-\Z}$, all initial-time correlators
 $\bk{\S_{\v m}}_{t=0}$ vanish unless $m_+=m_-=0$ (see Appendix
@@
 
-As a final example of collective-spin physics that is numerically
-accessible via the TST expansion of Heisenberg operators, we consider
-the calculation of collective-spin two-time correlation functions.  In
-particular, we consider the effect of decoherence on the short-time
-behavior of the two-time connected correlator
-\begin{align}
-  C\p{t}
-  \equiv \f1S\sp{\bk{S_+\p{t}S_-\p{0}}-\bk{S_+\p{t}}\bk{S_-\p{0}}}
-  \label{eq:two_time}
-\end{align}
-in the context of the squeezing models in Section \ref{sec:squeezing}.
-In an equilibrium setting, correlation functions of this sort contain
-information about the linear response of Heisenberg operators to
-perturbations of a system[CITE]\cite{CITE}; in a non-equilibrium
-setting, they contribute to short-time linear response (see Appendix
-\ref{sec:linear_response}).  Similar correlators have made appearances
-as order parameters that diagnose time-crystalline phases of
-matter\cite{tucker2018shattered}, and related multi-time correlators,
-known as out-of-time-ordered correlators, are commonly examined for
-signatures of quantum chaos and information
-scrambling[CITE]\cite{CITE}.
-
 \begin{figure*}
   \centering
-  \includegraphics{Sp_Sm.pdf}
+  \includegraphics{two_time.pdf}
   \caption{Real (solid lines) and imaginary (dashed lines) parts of
@@
 
+As a final example of collective-spin physics that is numerically
+accessible via the TST expansion of Heisenberg operators, we consider
+the calculation of collective-spin two-time correlation functions.  In
+particular, we consider the effect of decoherence on the short-time
+behavior of the two-time connected correlator
+\begin{align}
+  C\p{t}
+  \equiv \f1S\sp{\bk{S_+\p{t}S_-\p{0}}-\bk{S_+\p{t}}\bk{S_-\p{0}}}
+  \label{eq:two_time}
+\end{align}
+in the context of the squeezing models in Section \ref{sec:squeezing}.
+In an equilibrium setting, correlation functions of this sort contain
+information about the linear response of Heisenberg operators to
+perturbations of a system; in a non-equilibrium setting, they
+contribute to short-time linear response (see Appendix
+\ref{sec:linear_response}).  Similar correlators have made appearances
+as order parameters that diagnose time-crystalline phases of
+matter\cite{tucker2018shattered}, and related multi-time correlators,
+known as out-of-time-ordered correlators, are commonly examined for
+signatures of quantum chaos and information
+scrambling\cite{maldacena2016bound}.
+
 Figure \ref{fig:two_time} shows the behavior of $C\p{t}$ for $N=10^4$
@@
 straightforward to extend to multi-time and out-of-time-ordered
-correlators (OTOCs), as well as equilibrium settings.
+correlators, as well as equilibrium settings.
 
@@
 may be used to benchmark the short-time effects of decoherence, or
-compute quantities such as out-of-time-ordered correlators (OTOCs)
-that diagnose the onset of quantum chaos and information scrambling.
-
-\section{Acknowledgments}
-
+compute quantities such as out-of-time-ordered correlators that
+diagnose the onset of quantum chaos and information scrambling.
+
+\begin{acknowledgments}
   We acknowledge helpful discussions with Robert Lewis-Swan and Kris
-Tucker.  This work is supported by the Air Force Office of Scientific
-Research (AFOSR) grant FA9550-18-1-0319; the AFOSR Multidisciplinary
-University Research Initiative (MURI) grant; the Defense Advanced
-Research Projects Agency (DARPA) and Army Research Office (ARO) grant
-W911NF-16-1-0576; the National Science Foundation (NSF) grant
-PHY-1820885; JILA-NSF grant PFC-173400; and the National Institute of
-Standards and Technology (NIST).
-
-\newpage
+  Tucker.  This work is supported by the Air Force Office of
+  Scientific Research (AFOSR) grant FA9550-18-1-0319; the AFOSR
+  Multidisciplinary University Research Initiative (MURI) grant; the
+  Defense Advanced Research Projects Agency (DARPA) and Army Research
+  Office (ARO) grant W911NF-16-1-0576; the National Science Foundation
+  (NSF) grant PHY-1820885; JILA-NSF grant PFC-173400; and the National
+  Institute of Standards and Technology (NIST).
+\end{acknowledgments}
+
 \onecolumngrid
@@
 prove convenient for the calculations in this section\footnote{In
-  retrospect, it would likely have been nicer to have all
-  $s_\mu^{(j)}$ operators on the right throughout these calculations,
-  due to the enhanced symmetry expressions would have with respect to
-  Hermitian conjugation.  In any case, we provide the final result of
-  this section in both ordering conventions, and therefore feel no
-  need to reproduce these calculations with a different ordering of
-  spin operators.}.  This choice of ordering compels us to expand
+  retrospect, it may have been nicer to push all $s_\mu^{(j)}$
+  operators to the right throughout these calculations, due to the
+  enhanced symmetry expressions would have with respect to Hermitian
+  conjugation.  In any case, we provide the final result of this
+  section in both ordering conventions, and therefore feel no need to
+  reproduce these calculations with a different ordering of spin
+  operators.}.  This choice of ordering compels us to expand
 \begin{align}

### 2019-04-13 soc_squeezing/scripts/plot_squeezing.py
added one word to paper; changed colors in plot_squeezing.py; added notebook file to make two-time correlator plots for collective spin paper

@@ -51,6 +51,7 @@ dec_rates_strong = [ (100,)*3, (0,)*3 ]
 
 OAT, TAT, TNT = "OAT", "TAT", "TNT"
 methods = [ OAT, TAT, TNT ]
+colors = { OAT: "#4E79A7", TAT: "#F28E2B", TNT : "#E15759" }
 
 # construct Hamiltonians in (z,x,y) format, assuming an initial state in +X
 h_vec = { OAT : { (2,0,0) : 1 },
@@ -257,16 +258,16 @@ plt.figure(figsize = figsize)
 color = {}
 for method in methods:
     line, = plt.semilogy(time_sqz_C_exact[method][0],
-                         time_sqz_C_exact[method][1], label = method)
-    color[method] = line.get_color()
+                         time_sqz_C_exact[method][1],
+                         color = colors[method], label = method)
     positive_vals = positive(time_sqz_C_trunc[method][1])
     plt.semilogy(time_sqz_C_trunc[method][0][:positive_vals],
                  time_sqz_C_trunc[method][1][:positive_vals],
-                 "--", linewidth = trunc_width, color = darken_color(color[method]))
+                 "--", linewidth = trunc_width, color = darken_color(colors[method]))
     if positive_vals < len(times):
         plt.semilogy([time_sqz_C_trunc[method][0][positive_vals-1]],
                      [time_sqz_C_trunc[method][1][positive_vals-1]],
-                     "o", color = darken_color(color[method]))
+                     "o", color = darken_color(colors[method]))
 
 plt.xlabel(r"$N\chi t$")
 plt.ylabel(r"$\xi^2$")
@@ -286,15 +287,15 @@ for method in methods:
     linestyle = "-" if method == OAT else "."
     plt.semilogy(time_sqz_D_exact[method][0],
                  time_sqz_D_exact[method][1], linestyle, label = method,
-                 color = color[method], markersize = trajectory_marker_size)
+                 color = colors[method], markersize = trajectory_marker_size)
     positive_vals = positive(time_sqz_D_trunc[method][1])
     plt.semilogy(time_sqz_D_trunc[method][0][:positive_vals],
                  time_sqz_D_trunc[method][1][:positive_vals],
-                 "--", linewidth = trunc_width, color = darken_color(color[method]))
+                 "--", linewidth = trunc_width, color = darken_color(colors[method]))
     if positive_vals < len(times):
[…]

### 2019-04-15 soc_squeezing/collective_paper/collective_spin.tex
added one word to paper; changed colors in plot_squeezing.py; added notebook file to make two-time correlator plots for collective spin paper
mostly 'cosmetic' changes / minor wording changes from AMR

@@
 \newcommand{\f}[2]{\dfrac{#1}{#2}} % shorthand for fractions
-\newcommand{\p}[1]{\left(#1\right)} % parenthesis
-\renewcommand{\sp}[1]{\left[#1\right]} % square parenthesis
-\renewcommand{\set}[1]{\left\{#1\right\}} % curly parenthesis
+\newcommand{\p}[1]{\left( #1 \right)} % parenthesis
+\renewcommand{\sp}[1]{\left[ #1 \right]} % square parenthesis
 \renewcommand{\v}{\bm} % bold vectors
-\newcommand{\uv}[1]{\v{\hat{#1}}} % unit vectors
-\renewcommand{\c}{\cdot} % inner product
-\newcommand{\bk}{\Braket} % shorthand for braket notation
-
+\newcommand{\uv}[1]{\bm{\hat{#1}}} % unit vectors
+
+\newcommand{\bk}{\braket} % shorthand for braket notation
+
+\newcommand{\A}{\mathcal{A}}
 \newcommand{\C}{\mathcal{C}}
@@
 \newcommand{\I}{\mathcal{I}}
-\newcommand{\J}{\mathcal{J}}
 \renewcommand{\O}{\mathcal{O}}
@@
 
-\usepackage[inline]{enumitem} % for inline enumeration
+\renewcommand{\a}{\alpha} % free index
+\renewcommand{\b}{\beta} % free index
 
@@
 \graphicspath{{./figures/}} % set path for all figures
+
+\usepackage[inline]{enumitem} % for inline enumeration
 
@@
   We present a new method to compute short-time expectation values in
-  collective spin systems with permutationally-symmetric decoherence.
-  Our method is based on a Taylor expansion of a formal solution to
-  the equations of motion for Heisenberg operators.  This expansion
-  can be truncated at finite order to obtain nearly exact results at
-  short times that are relevant for metrological applications such as
-  spin squeezing.  In order to evaluate the expansion for Heisenberg
-  operators, we compute the structure constants of a collective SU(2)
-  operator algebra.  We demonstrate the utility of our method by
-  computing spin squeezing and two-time correlation functions in
-  strong-decoherence regimes that are inaccessible to existing
-  numerical methods.  Our method is straightforward to generalize to
-  the case of a collective spin coupled to bosonic modes, as can occur
-  in trapped ion and optical cavity experiments, and may be used to
-  compute quantities such as out-of-time-ordered correlators that
-  contain short-time signatures of quantum chaos and information
-  scrambling.
+  large ensembles of collective spin systems with generic Markovian
+  decoherence.  Our method is based on a Taylor expansion of a formal
+  solution to the equations of motion for Heisenberg operators.  This
+  expansion can be truncated at finite order to obtain virtually exact
+  results at short times that are relevant for metrological
+  applications such as spin squeezing.  In order to evaluate the
+  expansion for Heisenberg operators, we compute the relevant
+  structure constants of a collective operator algebra.  We
+  demonstrate the utility of our method by computing spin squeezing
+  and two-time correlation functions in strong-decoherence regimes
+  that are inaccessible to existing numerical methods.  Our method can
+  be straightforwardly generalized to the case of a collective spin
+  coupled to bosonic modes, relevant for trapped ion and cavity QED
+  experiments, and may be used to compute quantities such as
+  out-of-time-ordered correlators that can feature short-time
+  signatures of quantum chaos and information scrambling.
 \end{abstract}
@@
 long as measurements do not distinguish between constituent particles,
-even non-uniform systems can be effectively described by a uniform
+even non-uniform systems may be effectively described by a uniform
 model with renormalized parameters\cite{hu2015entangled}.
@@
 size $N$, thereby admitting efficient classical simulation of its
-dynamics.  Decoherence will generally violate total spin conservation
-and require the use of density operators, increasing the dimension of
+dynamics.  Decoherence generally violates total spin conservation and
+requires the use of density operators, increasing the dimension of
 accessible state space to $\sim N^3$\cite{hartmann2016generalized,
-  xu2013simulating} and allowing for exact simulations of
-$N\lesssim100$ particles.  If decoherence is sufficiently weak,
-dynamics can be numerically solvable for $N\lesssim10^5$ particles via
-``quantum trajectory'' Monte Carlo methods\cite{plenio1998quantumjump,
-  zhang2018montecarlo} (also known as ``quantum jump'' or ``Monte
-Carlo wavefunction'' methods) that can reproduce all expectation
-values of interest.  When decoherence is strong, however, or when the
-number of jump operators grows extensively with system size (e.g.~for
-single-spin decay), these Monte Carlo methods can take a prohibitively
-long time to converge, as simulations become dominated by incoherent
-jumps that generate large numbers of distinct quantum trajectories
-that need to be averaged over to compute expectation values.
-
-In this work, we present a new method to simulate short-time dynamics
-of collective spin systems with permutationally-symmetric sources of
-decoherence.  This method is based on a short-time expansion of exact
-solutions to the equations of motion for Heisenberg operators.
-Evaluating this expansion requires using the structure constants of a
-collective spin operator algebra; the calculation of these structure
-constants (in Appendices \ref{sec:identities}--\ref{sec:prod_general})
-is one of the main technical results of this work, which we hope will
-empower both analytical and numerical studies of collective spin
-systems in the future.  We benchmark our method against both exact and
-quantum trajectory Monte Carlo computations of spin squeezing in
-accessible parameter regimes, highlighting both advantages and
-limitations of our short-time expansion.  Finally, we showcase an
-application of our method with computations that are inaccessible to
-other numerical methods.
-
-[More text about the content of this paper.]
+  xu2013simulating}.  In this case, exact simulations can be carried
+out for $N\lesssim100$ particles.  If decoherence is sufficiently
+weak, dynamics can be numerically solvable for $N\lesssim10^5$
+particles via ``quantum trajectory'' Monte Carlo
+methods\cite{plenio1998quantumjump, zhang2018montecarlo} (also known
+as ``quantum jump'' or ``Monte Carlo wavefunction'' methods) that can
+reproduce all expectation values of interest.  When decoherence is
+strong, however, or when the number of jump operators grows
+extensively with system size (e.g.~for single-spin decay), these Monte
+Carlo methods can take a prohibitively long time to converge, as
+simulations become dominated by incoherent jumps that generate large
+numbers of distinct quantum trajectories that need to be averaged in
+order to compute expectation values.
+
+In this work, we present a new method to compute short-time dynamics
+of collective spin systems with generic Markovian decoherence.  The
+only restriction on decoherence is that, like the coherent collective
+dynamics, it must act identically on all constituent particles.  This
+method is based on a short-time expansion of exact solutions to the
+equations of motion for Heisenberg operators.  Evaluating this
+expansion requires knowing the structure constants of a collective
+spin operator algebra; the calculation of these structure constants
+(in Appendices \ref{sec:identities}--\ref{sec:prod_general}) is one of
+the main technical results of this work, which we hope will empower
+both analytical and numerical studies of collective spin systems in
+the future.  We benchmark our method against both exact and quantum
+trajectory Monte Carlo computations of spin squeezing in accessible
+parameter regimes, highlighting both advantages and limitations of our
+short-time expansion.  Finally, we showcase applications of our method
+with computations that are inaccessible to other numerical methods.
 
@@
 distinct spin-1/2 particles.  Defining individual spin-1/2 operators
-$s_{n=\x,\y,\z}\equiv\sigma_n/2$ and
-$s_\pm\equiv s_\x\pm is_\y=\sigma_\pm$ with Pauli operators
-$\sigma_n$, we denote an operator that acts with $s_n$ on the spin
-indexed by $j$ and trivially (i.e.~with the identity $\1$) on all
-other spins by $s_n^{(j)}$.  We then define the collective spin
-operators $S_n\equiv\sum_{j=1}^Ns_n^{(j)}$ for
-$n\in\set{\x,\y,\z,+,-}$.  Identifying the set $\set{\S_{\v m}}$ as a
-basis for all collective spin operators, with
+$\hat s_{\a=\x,\y,\z}\equiv\hat\sigma_\a/2$ and
+$\hat s_\pm\equiv\hat s_\x\pm i\hat s_\y=\hat\sigma_\pm$ with Pauli
+operators $\hat\sigma_\a$, we denote an operator that acts with
+$\hat s_\a$ on the spin indexed by $j$ and trivially (i.e.~with the
+identity $\hat\1$) on all other spins by $\hat s_\a^{(j)}$.  We then
+define the collective spin operators
+$\hat S_\a\equiv\sum_{j=1}^N\hat s_\a^{(j)}$ for
+$\a\in\set{\x,\y,\z,+,-}$.  Identifying the set $\set{\hat\S_{\v m}}$
+as a basis for all collective spin operators, with
 $\v m\equiv\p{m_+,m_\z,m_-}\in\N_0^3$ and
-$\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$, we can expand all
-collective spin operators in the form
-\begin{align}
-  \O = \sum_{\v m} \O_{\v m} \S_{\v m}.
-\end{align}
-The time evolution of a Heisenberg operator $\O\p{t}$ is then
-determined by
-\begin{align}
-  \f{d}{dt} \O\p{t}
-  = T \O\p{t}
-  = \sum_{\v m, \v n} \S_{\v m} T_{\v m\v n} \O_{\v n}\p{t}
+$\hat\S_{\v m}\equiv \hat S_+^{m_+} \hat S_\z^{m_\z} \hat S_-^{m_-}$,
+we can expand any collective spin operator $\hat\O$ in the form
+\begin{align}
+  \hat\O = \sum_{\v m} \O_{\v m} \hat\S_{\v m}
+\end{align}
+with scalar coefficients $\O_{\v m}\in\mathbb{C}$.  If $\hat\O$ is
+self-adjoint, for example, then $\O_{\v m}^*=\O_{\v m^*}$ with
+$\v m^*\equiv\p{m_-,m_\z,m_+}$.  The corresponding Heisenberg operator
+is then $\hat\O\p{t}=\sum_{\v m}\O_{\v m}\p{t}\hat\S_{\v m}$, with
+time-dependent coefficients $\O_{\v n}\p{t}$ for time-independent
+(Schr\"odinger) operators $\hat\S_{\v m}$.  The time evolution of
+Heisenberg operators is determined by
+\begin{align}
+  \f{d}{dt} \hat\O\p{t}
+  = \check T \hat\O\p{t}
+  = \sum_{\v m, \v n} \hat\S_{\v m} T_{\v m\v n} \O_{\v n}\p{t}
   \label{eq:EOM}
 \end{align}
-for a time derivative (super-)operator $T$ with matrix elements
-$T_{\v m\v n}$ defined by
-\begin{align}
-  T \S_{\v n} \equiv i \sp{H, \S_{\v n}}_-
-  + \sum_\J \gamma_\J \D\p{\J} \S_{\v n}
-  = \sum_{\v m} \S_{\v m} T_{\v m\v n},
+for a time derivative (super-)operator $\check T$ with matrix elements
+$T_{\v m\v n}\in\mathbb{C}$ defined by
+\begin{align}
+  \check T \hat\S_{\v n} \equiv i \sp{\hat H, \hat\S_{\v n}}_-
+  + \sum_\A \gamma_\A \check\D\p{\A} \hat\S_{\v n}
+  = \sum_{\v m} \hat\S_{\v m} T_{\v m\v n},
   \label{eq:time_deriv}
 \end{align}
-where $\sp{X,Y}_\pm\equiv XY\pm YX$; $H$ is the collective-spin
-Hamiltonian; $\J$ is a set of jump operators with a corresponding
-decoherence rate $\gamma_\J$; and is $\D\p{\J}$ is a
-Heisenberg-picture dissipator, or Lindblad superoperator, defined by
-\begin{align}
-  \D\p{\J} \O
-  \equiv \sum_{J\in\J} \p{J^\dag \O J - \f12\sp{J^\dag J,\O}_+}.
-\end{align}
-Decoherence via uncorrelated decay of individual spins at a rate
-$\gamma_-$, for example, would be described by the set of jump
-operators $\J_-=\set{s_-^{(j)}:j=1,2,\cdots,N}$ with
-$\gamma_{\J_-}=\gamma_-$.  The commutator in Eq.~\eqref{eq:time_deriv}
-can be computed by expanding the product
-$\S_{\v\ell}\S_{\v m}=\sum_{\v n}f_{\v\ell\v m\v n}\S_{\v n}$ with
-structure constants $f_{\v\ell\v m\v n}$ that are worked out in
-Appendices \ref{sec:identities}--\ref{sec:prod_general}, and the
-effects of decoherence from jump operators (i.e.~elements of $\J$) of
-the form $\gamma^{(j)}=\sum_n\gamma_ns_n^{(j)}$ and
-$\Gamma=\sum_n\Gamma_nS_n$ are worked out in Appendices
+where $\sp{X,Y}_\pm\equiv XY\pm YX$; $\hat H$ is the collective spin
+Hamiltonian; $\A$ is a set of jump operators with a corresponding
+decoherence rate $\gamma_\A$; and $\check\D$ is a Heisenberg-picture
+dissipator, or Lindblad superoperator, defined by
+\begin{align}
+  \check\D\p{\A} \hat \O
+  \equiv \sum_{\hat A\in\A} \p{\hat A^\dag \hat \O \hat A
+    - \f12\sp{\hat A^\dag \hat A, \hat \O}_+}.
+\end{align}
+Decoherence via uncorrelated decay of individual spins, for example,
+would be described by the set of jump operators
+$\A_-=\set{\hat s_-^{(j)}:j=1,2,\cdots,N}$.  The commutator in
+Eq.~\eqref{eq:time_deriv} can be computed by expanding the product
+$\hat\S_{\v\ell}\hat\S_{\v m} = \sum_{\v n} f_{\v\ell\v m\v
+  n}\hat\S_{\v n}$ with structure constants
+$f_{\v\ell\v m\v n}\in\mathbb{R}$ that we work out in Appendices
+\ref{sec:identities}--\ref{sec:prod_general}, and the effects of
+decoherence from jump operators (i.e.~elements of $\A$) of the form
+$\hat g^{(j)} = \sum_\a g_\a \hat s_\a^{(j)}$ and
+$\hat G = \sum_\a G_\a \hat S_\a$ are worked out in Appendices
 \ref{sec:sandwich_single}--\ref{sec:decoherence_collective}.  We
@@
 sufficient to compute matrix elements $T_{\v m\v n}$ of the time
-derivative operator $T$ in Eq.~\eqref{eq:time_deriv} in most cases of
-practical interest.
-
-The time derivative operator $T$ will generally couple spin operators
-$\S_{\v n}$ to spin operators $\S_{\v m}$ with higher ``weight'',
-i.e.~with $\abs{\v m}>\abs{\v n}$, where
-$\abs{\v\ell}\equiv\sum_\alpha\ell_\alpha$.  The growth of operator
-weight signifies the growth of many-body correlations.  In practice,
-keeping track of this growth will eventually require more
-computational resources than are available, meaning we must somehow
-truncate our equations of motion.  The simplest truncation strategy
-would be to take
-\begin{align}
-  \f{d}{dt} \O\p{t}
-  \to \sum_{w\p{\v m}<W} \S_{\v m}
-  \sum_{\v n} T_{\v m\v n} \O_{\v n}\p{t}
+derivative operator $\check T$ in Eq.~\eqref{eq:time_deriv} in most
+cases of practical interest.
+
+The time derivative operator $\check T$ will generally couple spin
+operators $\hat\S_{\v n}$ to spin operators $\hat\S_{\v m}$ with
+higher ``weight'', i.e.~with $\abs{\v m}>\abs{\v n}$, where
+$\abs{\v\ell}\equiv\sum_\a\ell_\a$.  The growth of operator weight
+signifies the growth of many-body correlations.  In practice, keeping
+track of this growth will eventually require more computational
+resources than are available, meaning we must somehow truncate our
+equations of motion.  The simplest truncation strategy would be to
+take
+\begin{align}
+  \f{d}{dt} \hat\O\p{t}
+  \to \sum_{w\p{\v m}<W} \hat\S_{\v m}
+  \sum_{\v n} T_{\v m\v n} \hat\O_{\v n}\p{t}
   \label{eq:weight_truncation}
@@
 \item simulation results can only be trusted up to the time at which
-  the initial value of operators $\S_{\v m}$ with weight
+  the initial value of operators $\hat\S_{\v m}$ with weight
   $w\p{\v m}\ge W$ have a non-negligible contribution to expectation
@@
 therefore devise an alternate truncation strategy built around
-limitation \ref{pt:limitation}.
-
-We can expand Heisenberg operators $\O\p{t}$ in a Taylor series about
-the time $t=0$ to write
-\begin{align}
-  \O\p{t}
-  = e^{t T} \O\p{0}
+limitation \ref*{pt:limitation}.
+
+We can formally expand Heisenberg operators $\hat\O\p{t}$ in a Taylor
+series about the time $t=0$ to write
+\begin{align}
+  \hat\O\p{t}
+  = e^{t\check T} \hat \O\p{0}
   = \sum_{k\ge0} \f{t^k}{k!}
-  \sum_{\v m, \v n} \S_{\v m} T^k_{\v m\v n} \O_{\v n}\p{0},
+  \sum_{\v m, \v n} \hat\S_{\v m} T^k_{\v m\v n} \O_{\v n}\p{0},
   \label{eq:time_series}
@@
 where the matrix elements $T^k_{\v m\v n}$ of the $k$-th time
-derivative operator $T^k$ are
+derivative operator are
 \begin{align}
@@
 \begin{align}
-  \O\p{t}
+  \hat\O\p{t}
   \to \sum_{k=0}^M \f{t^k}{k!}
-  \sum_{\v m, \v n} \S_{\v m} T^k_{\v m\v n} \O_{\v n}\p{0}.
+  \sum_{\v m, \v n} \hat\S_{\v m} T^k_{\v m\v n} \O_{\v n}\p{0}.
   \label{eq:order_truncation}
@@
 Eq.~\eqref{eq:order_truncation} tell us which operators are relevant
-for computing the expectation value $\bk{\O\p{t}}$ to a fixed order
-$M$.  Note that using the relation $\S_{\v m}^\dag=\S_{\v m^*}$ for
-$\v m^*\equiv\p{m_-,m_\z,m_+}$, which by Hermitian conjugation of
-Eq.~\eqref{eq:EOM} also implies that
+for computing the expectation value $\bk{\hat\O\p{t}}$ to a fixed
+order $M$.  Note that using the relation
+$\hat\S_{\v m}^\dag=\hat\S_{\v m^*}$, which by Hermitian conjugation
+of Eq.~\eqref{eq:EOM} also implies that
 $T_{\v m^*\v n^*}=T_{\v m\v n}^*$, can cut both the number of
-initial-time expectation values $\bk{\S_{\v m}}_{t=0}$ and the number
-qof matrix elements $T_{\v m\v n}$ that we may need to explicitly
-compute roughly in half.
+initial-time expectation values $\bk{\hat\S_{\v m}}_{t=0}$ and the
+number qof matrix elements $T_{\v m\v n}$ that we may need to
+explicitly compute roughly in half.
 
@@
 with limited computational resources, such that the TST expansion in
-Eq.~\eqref{eq:order_truncation} only allows computation of the
-expectation value $\bk{\O\p{t}}$ to a maximal time $t_{M_{\t{max}}}$.
-As we will see in the following section, $M_{\t{max}}=35$ will suffice
-for the computation of collective spin correlators up to times that
-are relevant e.g.~for spin squeezing protocols.
+Eq.~\eqref{eq:order_truncation} only allows for computation of the
+expectation value $\bk{\hat\O\p{t}}$ to a maximal time
+$t_{M_{\t{max}}}$.  As we will see in the following section,
+$M_{\t{max}}=35$ will suffice for the computation of collective spin
+correlators up to times that are relevant e.g.~for spin squeezing
+protocols.
 
@@
 \begin{align}
-  H_{\t{OAT}} &= \chi S_\z^2, \label{eq:OAT} \\
-  H_{\t{TAT}} &= \f{\chi}{3} \p{S_\z^2 - S_\y^2}, \label{eq:TAT} \\
-  H_{\t{TNT}} &= \chi S_\z^2 + \Omega S_\x, \label{eq:TNT}
+  H_{\t{OAT}} &= \chi \hat S_\z^2, \label{eq:OAT} \\
+  H_{\t{TAT}}
+  &= \f{\chi}{3} \p{\hat S_\z^2 - \hat S_\y^2}, \label{eq:TAT} \\
+  H_{\t{TNT}} &= \chi \hat S_\z^2 + \Omega \hat S_\x, \label{eq:TNT}
 \end{align}
@@
 initial product state $\ket{\X}\propto\p{\ket\up+\ket\dn}^{\otimes N}$
-with $S_\x\ket{\X}=S\ket{\X}$.  Our measure of spin squeezing is the
-directionally-unbiased Ramsey squeezing parameter determined by the
-maximal gain in resolution $\Delta\phi$ of a phase angle $\phi$ over
-that achieved by any spin-polarized product state
+with $\hat S_\x\ket{\X}=S\ket{\X}$.  Our measure of spin squeezing is
+the directionally-unbiased Ramsey squeezing parameter determined by
+the maximal gain in resolution $\Delta\phi$ of a phase angle $\phi$
+over that achieved by any spin-polarized product state
 (e.g.~$\ket{\X}$)\cite{wineland1992spin, ma2011quantum},
@@
   \equiv \f{\p{\Delta\phi_{\t{min}}}^2}{\p{\Delta\phi_{\t{polarized}}}^2}
-  = \f{N}{\abs{\bk{\v S}}^2}
-  \min_{\uv n\perp\bk{\v S}} \bk{\p{\v S\c\uv n}^2},
+  = \f{N}{\lvert\braket{\uv S}\rvert^2}
+  \min_{\uv n\perp\bk{\uv S}} \Braket{\p{\uv S\cdot\uv n}^2},
   \label{eq:squeezing}
 \end{align}
-where $\v S\equiv\p{S_\x,S_\y,S_\z}$ is a collective spin vector, and
-the minimization is performed over all unit vectors $\uv n$ in the
-plane orthogonal to the mean spin vector $\bk{\v S}$.  This squeezing
-parameter is entirely determined by one- and two-spin correlators of
-the form $\bk{S_m}$ and $\bk{S_m S_n}$.  In the case of the coherent
-dynamics discussed here, these correlators are obtainable via exact
-simulations of quantum dynamics in the $\p{N+1}$-dimensional Dicke
-manifold of states $\set{\ket{S,m}}$ with net spin $S$ and spin
-projection $m$ onto the $z$ axis, i.e.~with
-$\bk{S,m|\v S^2|S,m}=S\p{S+1}$ and $\bk{S,m|S_\z|S,m}=m$ for
-$m\in\set{-S,-S+1,\cdots,S}$.  In the presence of
-permutationally-symmetric decoherence, these correlators are
-obtainable with the collective-spin quantum trajectory Monte Carlo
-method developed in ref.~\cite{zhang2018montecarlo}.  In this work,
-these exact and quantum trajectory simulations will be used to
-benchmark the TST expansion in Eq.~\eqref{eq:order_truncation}.
+where $\uv S\equiv\p{\hat S_\x,\hat S_\y,\hat S_\z}$ is a collective
+spin operator-valued vector, and the minimization is performed over
+all unit vectors $\uv n$ in the plane orthogonal to the mean spin
+vector $\bk{\uv S}$.  This squeezing parameter is entirely determined
+by one- and two-spin correlators of the form $\bk{\hat S_\a}$ and
+$\bk{\hat S_\a \hat S_\b}$.  For the unitary dynamics discussed in this
+work, these correlators are obtainable via exact simulations of
+quantum dynamics in the $\p{N+1}$-dimensional Dicke manifold of states
+$\set{\ket{S,m}}$ with net spin $S$ and spin projection $m$ onto the
+$z$ axis, i.e.~with $\bk{S,m|\uv S^2|S,m}=S\p{S+1}$ and
+$\bk{S,m|\hat S_\z|S,m}=m$ for $m\in\set{-S,-S+1,\cdots,S}$.  In the
+presence of single-spin or collective decoherence, meanwhile, these
+correlators are obtainable with the collective-spin quantum trajectory
+Monte Carlo method developed in ref.~\cite{zhang2018montecarlo}.  In
+this work, these exact and quantum trajectory simulations will be used
+to benchmark the TST expansion in Eq.~\eqref{eq:order_truncation}.
 
@@
   \centering
-  \subfloat[Squeezing with coherent dynamics
+  \subfloat[Squeezing with unitary dynamics
   \label{fig:squeezing_coherent}]
@@
   \caption{Spin squeezing of $N=10^4$ spins initially in $\ket\X$
-    under \sref{fig:squeezing_coherent} coherent and
-    \sref{fig:squeezing_incoherent} incoherent evolution, computed
+    under \sref{fig:squeezing_coherent} unitary and
+    \sref{fig:squeezing_incoherent} non-unitary dynamics, computed
     using exact methods (solid lines), quantum trajectory simulations
@@
 Eq.~\eqref{eq:order_truncation} with $M=35$.  Squeezing is shown for
-both coherent dynamics (Figure \ref{fig:squeezing_coherent}), as well
-as dynamics in the presence of spontaneous decay, excitation, and
-dephasing of individual spins at rates $\chi$ (Figure
+both unitary dynamics (Figure \ref{fig:squeezing_coherent}), as well
+as non-unitary dynamics in the presence of spontaneous decay,
+excitation, and dephasing of individual spins at rates $\chi$ (Figure
 \ref{fig:squeezing_incoherent}), respectively described by the sets of
-jump operators $\J_n=\set{s_n^{(j)}}$ with corresponding decoherence
-rates $\gamma_n=\chi$ for $n\in\set{-,+,\z}$.  The results shown in
-Figure \ref{fig:benchmarking} were computed in a rotated basis with
-$\p{s_\z,s_\x}\to\p{s_\x,-s_\z}$ and
+jump operators $\A_\a=\set{\hat s_\a^{(j)}}$ with corresponding
+decoherence rates $\gamma_\a=\chi$ for $\a\in\set{-,+,\z}$.  The
+results shown in Figure \ref{fig:benchmarking} were computed in a
+rotated basis with
+$\p{\hat s_\z,\hat s_\x}\to\p{\hat s_\x,-\hat s_\z}$ and
 $\ket\X\to\ket{-\Z}\equiv\ket\dn^{\otimes N}$ (together with
@@
 departure occurs because the TST expansion neglects high-weight
-(i.e.~large-$\abs{\v m}$) operators $\S_{\v m}$ whose contributions to
-a Heisenberg operator $\O\p{t}$ of interest eventually become
-non-negligible.  For sufficiently large times $t$, these individual
-contributions may generally be large compared to the value of
-$\bk{\O\p{t}}$, in which case their truncation will yield non-physical
-results.  This breakdown mechanism is also the reason for prolonged
-agreement between the TST expansion and benchmarking methods when
-squeezing the initial state $\ket{-\Z}$ rather than $\ket{\X}$: for
-initial states $\ket{-\Z}$, all initial-time correlators
-$\bk{\S_{\v m}}_{t=0}$ vanish unless $m_+=m_-=0$ (see Appendix
+(i.e.~large-$\abs{\v m}$) operators $\hat\S_{\v m}$ whose
+contributions to a Heisenberg operator $\hat\O\p{t}$ of interest
+eventually become non-negligible.  For sufficiently large times $t$,
+these individual contributions may generally be large compared to the
+value of $\bk{\hat\O\p{t}}$, in which case their truncation will yield
+non-physical results.  This breakdown mechanism is also the reason for
+prolonged agreement between the TST expansion and benchmarking methods
+when squeezing the initial state $\ket{-\Z}$ rather than $\ket{\X}$:
+for initial states $\ket{-\Z}$, all initial-time correlators
+$\bk{\hat\S_{\v m}}_{t=0}$ vanish unless $m_+=m_-=0$ (see Appendix
 \ref{sec:initial_conditions}), so there is a substantially smaller
 number of neglected terms with non-zero contribution to the correlator
-$\bk{\O\p{t}}$.
+$\bk{\hat\O\p{t}}$.
 
@@
 in Figure \ref{fig:squeezing_incoherent}, meanwhile, take $\sim10$
-seconds to compute with a single CPU on similar computing hardware.
+seconds to compute with a single CPU on similar hardware.
+
+\begin{figure}
+  \centering
+  \includegraphics{decoherence_strong.pdf}
+  \caption{Spin squeezing of $N=10^4$ spins initially in $\ket\X$ with
+    spontaneous decay, excitation, and dephasing of individual spins
+    at rates $\gamma_-=\gamma_+=\gamma_\z=100\chi$.  Computed using
+    the TST expansion in Eq.~\eqref{eq:order_truncation} with $M=35$.
+    Solid circles mark the times at which the TST expansion gives an
+    unphysical result with $\xi^2<0$.}
+  \label{fig:decoherence_strong}
+\end{figure}
 
@@
 
-\begin{figure}
-  \centering
-  \includegraphics{decoherence_strong.pdf}
-  \caption{Spin squeezing of $N=10^4$ spins initially in $\ket\X$
-    under incoherent evolution spontaneous with decay, excitation, and
-    dephasing of individual spins at rates
-    $\gamma_-=\gamma_+=\gamma_\z=100\chi$, computed using the TST
-    expansion in Eq.~\eqref{eq:order_truncation} with $M=35$.  Solid
-    circles mark the times at which the TST expansion gives an
-    unphysical result with $\xi^2<0$.}
-  \label{fig:decoherence_strong}
-\end{figure}
-
-\section{Two-time correlation functions}
-\label{sec:two_time}
-
 \begin{figure*}
@@
     evolution without decoherence (leftmost panel), as well as with
-    decoherence at rates $\gamma_n$ or $\Gamma_n$, respectively
-    corresponding to single-spin or collective jump operators
-    $s_n^{(j)}$ or $S_n$.  All axes in the sub-figures are shared.  At
-    short times, the correlations captured by $C\p{t}$ are suppressed
-    by single-spin decay ($\gamma_-$), collective decay ($\Gamma_-$),
-    and single-spin dephasing ($\gamma_\z$); these correlations are
-    relatively insensitive to collective dephasing ($\Gamma_\z$), and
-    can change sign or grow from spontaneous single-spin ($\gamma_+$)
-    or collective ($\Gamma_+$) excitation.}
+    decoherence at rates $\gamma_\a$ or $\Gamma_\a$,
+    respectively corresponding to single-spin or collective jump
+    operators $s_\a^{(j)}$ or $S_\a$.  All axes in the sub-figures are
+    shared.  At short times, the correlations captured by $C\p{t}$ are
+    suppressed by single-spin decay ($\gamma_-$), collective decay
+    ($\Gamma_-$), and single-spin dephasing ($\gamma_\z$); these
+    correlations are relatively insensitive to collective dephasing
+    ($\Gamma_\z$), and can change sign or grow from spontaneous
+    single-spin ($\gamma_+$) or collective ($\Gamma_+$) excitation.}
   \label{fig:two_time}
 \end{figure*}
+
+\section{Two-time correlation functions}
+\label{sec:two_time}
 
@@
 the calculation of collective-spin two-time correlation functions.  In
-particular, we consider the effect of decoherence on the short-time
+particular, we consider the effect of decoherence on short-time
 behavior of the two-time connected correlator
@@
   C\p{t}
-  \equiv \f1S\sp{\bk{S_+\p{t}S_-\p{0}}-\bk{S_+\p{t}}\bk{S_-\p{0}}}
+  \equiv \f1S\sp{\bk{\hat S_+\p{t} \hat S_-\p{0}}
+    - \bk{\hat S_+\p{t}}\bk{\hat S_-\p{0}}}
   \label{eq:two_time}
@@
 decoherence, but seems relatively insensitive to collective dephasing
-(with jump operator $S_\z$), and can qualitatively change behavior or
-grow in the presence of single-spin or collective excitation (with
-jump operators $s_+^{(j)}$ or $S_+$).  Figure \ref{fig:two_time}
-serves as an example for the type of quantity and behavior that is
-accessible at short times with the TST expansion.  This example is
-straightforward to extend to multi-time and out-of-time-ordered
-correlators, as well as equilibrium settings.
+(with jump operator $\hat S_\z$), and can qualitatively change
+behavior or grow in the presence of single-spin or collective
+excitation (with jump operators $\hat s_+^{(j)}$ or $\hat S_+$).
+Figure \ref{fig:two_time} serves as an example for the type of
+behavior that is accessible at short times with the TST expansion.
+This example is straightforward to extend to multi-time and
+out-of-time-ordered correlators, as well as equilibrium settings.
 
@@
   Defense Advanced Research Projects Agency (DARPA) and Army Research
-  Office (ARO) grant W911NF-16-1-0576; the National Science Foundation
-  (NSF) grant PHY-1820885; JILA-NSF grant PFC-173400; and the National
-  Institute of Standards and Technology (NIST).
+  Office (ARO) grant W911NF-16-1-0576 and W911NF-19-1-0210; the
+  National Science Foundation (NSF) grant PHY-1820885; JILA-NSF grant
+  PFC-173400; and the National Institute of Standards and Technology
+  (NIST).
 \end{acknowledgments}
@@
 for reference.  Note that despite the working definition of collective
-spin operators from $S_n=\sum_js_n^{(j)}$, the identities we will
+spin operators from $S_\a=\sum_js_\a^{(j)}$, the identities we will
 derive involving only collective spin operators apply just as well to
@@
 impose an ordering with all $s_\bmu^{(j)}$ operators on the left, and
-all $s_\z^{(j)}$ operators on the right, as such an ordering will
-prove convenient for the calculations in this section\footnote{In
+all $s_\z^{(j)}$ operators on the right.  Such an ordering will prove
+convenient for the calculations in this section\footnote{In
   retrospect, it may have been nicer to push all $s_\mu^{(j)}$
@@
 The (anti-)commutator of two ordered products of collective spin
-operators is then
+operators is then simply
 \begin{align}
@@
 single-spin operators together, and then evaluate the sum to arrive at
-an expression in terms of only collective spin operators.
+an expression involving only collective spin operators.
 
@@
 \begin{align}
-  \D\p{\gamma} \O
-  \equiv \D\p{\set{\gamma^{(j)}:j=1,2,\cdots,N}} \O
-  = \sum_j\p{{\gamma^{(j)}}^\dag \O \gamma^{(j)}
-    - \f12\sp{{\gamma^{(j)}}^\dag \gamma^{(j)}, \O}_+},
-\end{align}
-where $\gamma$ is an operator that acts on a single spin,
-$\gamma^{(j)}$ is an operator that acts with $\gamma$ on spin $j$ and
-trivially on all other spins, and $N$ is the total number of spins.
+  \D\p{g} \O
+  \equiv \D\p{\set{g^{(j)}:j=1,2,\cdots,N}} \O
+  = \sum_j\p{{g^{(j)}}^\dag \O g^{(j)}
+    - \f12\sp{{g^{(j)}}^\dag g^{(j)}, \O}_+},
+\end{align}
+where $g$ is an operator that acts on a single spin, $g^{(j)}$ is an
+operator that acts with $g$ on spin $j$ and trivially on all other
+spins, and $N$ is the total number of spins.
 
@@
 \begin{align}
-  \D\p{\gamma} \O
-  = \sum_j\p{{\gamma^{(j)}}^\dag \O \gamma^{(j)}
-    - \f12\sp{{\gamma^{(j)}}^\dag \gamma^{(j)}, \O}_+},
+  \D\p{g} \O
+  = \sum_j\p{{g^{(j)}}^\dag \O g^{(j)}
+    - \f12\sp{{g^{(j)}}^\dag g^{(j)}, \O}_+},
   &&
-  \gamma \equiv \gamma_\z s_\z + \gamma_+ s_+ + \gamma_- s_-.
+  g \equiv g_\z s_\z + g_+ s_+ + g_- s_-.
   \label{eq:D_general_single}
@@
 \begin{align}
-  \gamma^\dag \O \gamma
-  = \abs{\gamma_\z}^2 s_\z \O s_\z
-  + \sum_\mu \p{\abs{\gamma_\mu}^2 s_\bmu \O s_\mu
-    + \gamma_\bmu^* \gamma_\mu s_\mu \O s_\mu
-    + \gamma_\z^* \gamma_\mu s_\z \O s_\mu
-    + \gamma_\bmu^* \gamma_\z s_\mu \O s_\z},
+  g^\dag \O g
+  = \abs{g_\z}^2 s_\z \O s_\z
+  + \sum_\mu \p{\abs{g_\mu}^2 s_\bmu \O s_\mu
+    + g_\bmu^* g_\mu s_\mu \O s_\mu
+    + g_\z^* g_\mu s_\z \O s_\mu
+    + g_\bmu^* g_\z s_\mu \O s_\z},
 \end{align}
@@
 \begin{align}
-  \gamma^\dag \gamma
-  = \f14 \abs{\gamma_\z}^2
-  + \f12 \sum_\mu \sp{\abs{\gamma_\mu}^2 \p{1-2\mu s_\z}
-    + \mu \p{\gamma_\z^*\gamma_\mu - \gamma_\bmu^*\gamma_\z} s_\mu},
+  g^\dag g
+  = \f14 \abs{g_\z}^2
+  + \f12 \sum_\mu \sp{\abs{g_\mu}^2 \p{1-2\mu s_\z}
+    + \mu \p{g_\z^*g_\mu - g_\bmu^*g_\z} s_\mu},
 \end{align}
@@
 \begin{align}
-  \D\p{\gamma} \O
-  &= \sum_{X\in\set{\z,+,-}} \abs{\gamma_X}^2 \D\p{s_X} \O
+  \D\p{g} \O
+  &= \sum_{X\in\set{\z,+,-}} \abs{g_X}^2 \D\p{s_X} \O
   + \sum_{\mu,j}
-  \p{\gamma_\bmu^* \gamma_\mu s_\mu^{(j)} \O s_\mu^{(j)}
-    + \gamma_\z^* \gamma_\mu s_\z^{(j)} \O s_\mu^{(j)}
-    + \gamma_\bmu^* \gamma_\z s_\mu^{(j)} \O s_\z^{(j)}}
+  \p{g_\bmu^* g_\mu s_\mu^{(j)} \O s_\mu^{(j)}
+    + g_\z^* g_\mu s_\z^{(j)} \O s_\mu^{(j)}
+    + g_\bmu^* g_\z s_\mu^{(j)} \O s_\z^{(j)}}
   \notag \\
   &\qquad -\f14 \sum_\mu \mu
-  \p{\gamma_\z^*\gamma_\mu - \gamma_\bmu^*\gamma_\z} \sp{S_\mu, \O}_+.
+  \p{g_\z^*g_\mu - g_\bmu^*g_\z} \sp{S_\mu, \O}_+.
 \end{align}
@@
 \begin{align}
-  Q_{\ell mn;\mu}^{(\gamma)}
-  \equiv \gamma_\bmu^* \gamma_\mu P_{\ell mn;\mu}
-  + \p{\gamma_\z^* \gamma_\mu + \gamma_\bmu^* \gamma_\z}
+  Q_{\ell mn;\mu}^{(g)}
+  \equiv g_\bmu^* g_\mu P_{\ell mn;\mu}
+  + \p{g_\z^* g_\mu + g_\bmu^* g_\z}
   K_{\ell mn;\mu}
-  + \gamma_\z^* \gamma_\mu L_{\ell mn;\mu}
-  + \gamma_\bmu^* \gamma_\z M_{\ell mn;\mu},
+  + g_\z^* g_\mu L_{\ell mn;\mu}
+  + g_\bmu^* g_\z M_{\ell mn;\mu},
   \label{eq:Q_single}
@@
 \begin{align}
-  \D\p{\gamma} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
-  = \sum_{X\in\set{\z,+,-}} \abs{\gamma_X}^2
+  \D\p{g} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
+  = \sum_{X\in\set{\z,+,-}} \abs{g_X}^2
   \D\p{s_X} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
-  + Q_{\ell mn;\mu}^{(\gamma)} + \sp{Q_{nm\ell;\mu}^{(\gamma)}}^\dag.
+  + Q_{\ell mn;\mu}^{(g)} + \sp{Q_{nm\ell;\mu}^{(g)}}^\dag.
 \end{align}
@@
 \begin{align}
-  \D\p{\Gamma} \O
-  \equiv \D\p{\set{\Gamma}} \O
-  = \Gamma^\dag \O \Gamma - \f12\sp{\Gamma^\dag \Gamma, \O}_+,
-\end{align}
-where $\Gamma$ is a collective spin jump operator.
+  \D\p{G} \O
+  \equiv \D\p{\set{G}} \O
+  = G^\dag \O G - \f12\sp{G^\dag G, \O}_+,
+\end{align}
+where $G$ is a collective spin jump operator.
 
@@
 \begin{align}
-  \Gamma \equiv \Gamma_\z S_\z + \Gamma_+ S_+ + \Gamma_- S_-,
+  G \equiv G_\z S_\z + G_+ S_+ + G_- S_-,
 \end{align}
@@
 \begin{align}
-  \Gamma^\dag \O \Gamma
-  = \abs{\Gamma_\z}^2 S_\z \O S_\z
-  + \sum_\mu \p{\abs{\Gamma_\mu}^2 S_\bmu \O S_\mu
-    + \Gamma_\bmu^* \Gamma_\mu S_\mu \O S_\mu
-    + \Gamma_\z^* \Gamma_\mu S_\z \O S_\mu
-    + \Gamma_\bmu^* \Gamma_\z S_\mu \O S_\z},
+  G^\dag \O G
+  = \abs{G_\z}^2 S_\z \O S_\z
+  + \sum_\mu \p{\abs{G_\mu}^2 S_\bmu \O S_\mu
+    + G_\bmu^* G_\mu S_\mu \O S_\mu
+    + G_\z^* G_\mu S_\z \O S_\mu
+    + G_\bmu^* G_\z S_\mu \O S_\z},
 \end{align}
@@
 \begin{align}
-  \Gamma^\dag \Gamma
-  = \abs{\Gamma_\z}^2 S_\z^2
-  + \sum_\mu \p{\abs{\Gamma_\mu}^2 S_\bmu S_\mu
-    + \Gamma_\z^*\Gamma_\mu S_\z S_\mu
-    + \Gamma_\bmu^* \Gamma_\z S_\mu S_\z
-    + \Gamma_\bmu^* \Gamma_\mu S_\mu^2},
+  G^\dag G
+  = \abs{G_\z}^2 S_\z^2
+  + \sum_\mu \p{\abs{G_\mu}^2 S_\bmu S_\mu
+    + G_\z^*G_\mu S_\z S_\mu
+    + G_\bmu^* G_\z S_\mu S_\z
+    + G_\bmu^* G_\mu S_\mu^2},
 \end{align}
@@
 \begin{align}
-  \D\p{\Gamma} \O
-  &= \sum_{X\in\set{\z,+,-}} \abs{\Gamma_X}^2 \D\p{S_X} \O
-  + \sum_\mu \p{\Gamma_\bmu^* \Gamma_\mu S_\mu \O S_\mu
-    + \Gamma_\z^* \Gamma_\mu S_\z \O S_\mu
-    + \Gamma_\bmu^* \Gamma_\z S_\mu \O S_\z}
+  \D\p{G} \O
+  &= \sum_{X\in\set{\z,+,-}} \abs{G_X}^2 \D\p{S_X} \O
+  + \sum_\mu \p{G_\bmu^* G_\mu S_\mu \O S_\mu
+    + G_\z^* G_\mu S_\z \O S_\mu
+    + G_\bmu^* G_\z S_\mu \O S_\z}
   \notag \\
-  &\qquad -\f12 \sum_\mu\p{\Gamma_\bmu^* \Gamma_\mu \sp{S_\mu^2, \O}_+
-    + \Gamma_\z^*\Gamma_\mu \sp{S_\z S_\mu, \O}_+
-    + \Gamma_\bmu^* \Gamma_\z \sp{S_\mu S_\z, \O}_+}.
+  &\qquad -\f12 \sum_\mu\p{G_\bmu^* G_\mu \sp{S_\mu^2, \O}_+
+    + G_\z^*G_\mu \sp{S_\z S_\mu, \O}_+
+    + G_\bmu^* G_\z \sp{S_\mu S_\z, \O}_+}.
 \end{align}
@@
 \begin{align}
-  \Gamma_{\z,\mu}^{(\pm)}
-  &\equiv \f12\p{\Gamma_\z^* \Gamma_\mu \pm \Gamma_\bmu^* \Gamma_\z},
+  G_{\z,\mu}^{(\pm)}
+  &\equiv \f12\p{G_\z^* G_\mu \pm G_\bmu^* G_\z},
   \allowdisplaybreaks \\
-  \tilde L_{\ell mn;\mu}^{(\Gamma)}
-  &\equiv \mu \sp{\p{\ell-n+\f12} \Gamma_{\z,\mu}^{(+)}
-    + \p{\ell+\f12} \Gamma_{\z,\mu}^{(-)}}
+  \tilde L_{\ell mn;\mu}^{(G)}
+  &\equiv \mu \sp{\p{\ell-n+\f12} G_{\z,\mu}^{(+)}
+    + \p{\ell+\f12} G_{\z,\mu}^{(-)}}
   S_\mu^{\ell+1} \p{1+\mu S_\z}^m S_\bmu^n \notag \\
-  &\qquad -\mu \sp{\p{\ell-n+\f12} \Gamma_{\z,\mu}^{(+)}
-    + \p{n+\f12} \Gamma_{\z,\mu}^{(-)}}
+  &\qquad -\mu \sp{\p{\ell-n+\f12} G_{\z,\mu}^{(+)}
+    + \p{n+\f12} G_{\z,\mu}^{(-)}}
   S_\mu^{\ell+1} \p{\mu S_\z}^m S_\bmu^n \notag \\
-  &\qquad + \mu \Gamma_{\z,\mu}^{(-)}
+  &\qquad + \mu G_{\z,\mu}^{(-)}
   S_\mu^{\ell+1} \mu S_\z
@@
   \allowdisplaybreaks \\
-  \tilde M_{\ell mn;\mu}^{(\Gamma)}
-  &= -\mu n\p{n-1} \sp{\p{\ell-n+\f12} \Gamma_{\z,\mu}^{(+)}
-    + \p{\ell-\f12} \Gamma_{\z,\mu}^{(-)}}
+  \tilde M_{\ell mn;\mu}^{(G)}
+  &= -\mu n\p{n-1} \sp{\p{\ell-n+\f12} G_{\z,\mu}^{(+)}
+    + \p{\ell-\f12} G_{\z,\mu}^{(-)}}
   S_\mu^\ell \p{\mu S_\z}^m S_\bmu^{n-1} \notag \\
-  &\qquad - 2\mu n \sp{\p{\ell-n+\f12} \Gamma_{\z,\mu}^{(+)}
-    + \p{\ell+\f12n-1} \Gamma_{\z,\mu}^{(-)}}
+  &\qquad - 2\mu n \sp{\p{\ell-n+\f12} G_{\z,\mu}^{(+)}
+    + \p{\ell+\f12n-1} G_{\z,\mu}^{(-)}}
   S_\mu^\ell \p{\mu S_\z}^{m+1} S_\bmu^{n-1} \notag \\
-  &\qquad - 2\mu n \Gamma_{\z,\mu}^{(-)}
+  &\qquad - 2\mu n G_{\z,\mu}^{(-)}
   S_\mu^\ell \p{\mu S_\z}^{m+2} S_\bmu^{n-1},
@@
   \allowdisplaybreaks \\
-  \tilde Q_{\ell mn;\mu}^{(\Gamma)}
-  &\equiv \Gamma_\bmu^* \Gamma_\mu \tilde P_{\ell mn;\mu}
-  + \tilde L_{\ell mn;\mu}^{(\Gamma)}
-  + \tilde M_{\ell mn;\mu}^{(\Gamma)},
+  \tilde Q_{\ell mn;\mu}^{(G)}
+  &\equiv G_\bmu^* G_\mu \tilde P_{\ell mn;\mu}
+  + \tilde L_{\ell mn;\mu}^{(G)}
+  + \tilde M_{\ell mn;\mu}^{(G)},
 \end{align}
@@
 \begin{align}
-  \D\p{\Gamma} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
-  = \sum_{X\in\set{\z,+,-}} \abs{\Gamma_X}^2
+  \D\p{G} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
+  = \sum_{X\in\set{\z,+,-}} \abs{G_X}^2
   \D\p{S_X} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
-  + \tilde Q_{\ell mn;\mu}^{(\Gamma)}
-  + \sp{\tilde Q_{nm\ell;\mu}^{(\Gamma)}}^\dag.
+  + \tilde Q_{\ell mn;\mu}^{(G)}
+  + \sp{\tilde Q_{nm\ell;\mu}^{(G)}}^\dag.
 \end{align}
 Note that the sum $\ell+m+n$ for operators
-$S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$ grows by one if $\Gamma_\mu\ne0$
-or $\Gamma_\bmu\ne0$, and does not grow otherwise.
+$S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$ grows by one if $G_\mu\ne0$ or
+$G_\bmu\ne0$, and does not grow otherwise.
 
@@
 \begin{align}
-  \widetilde T = T + i\hat V
-\end{align}
-where $\hat V$ is a superoperator whose action on operators $\O$ is
+  \widetilde T = T + i\check V
+\end{align}
+where $\check V$ is a superoperator whose action on operators $\O$ is
 defined by
 \begin{align}
-  \hat V \O \equiv \sp{V,\O}_-.
+  \check V \O \equiv \sp{V,\O}_-.
 \end{align}
 Through quadratic order in the time $t$ and linear order in the
-perturbation $\hat V$, we can say that
+perturbation $\check V$, we can say that
 \begin{align}
   e^{t\widetilde T}
-  \approx \f12\sp{e^{tT}, e^{it\hat V}}_+
-  \approx e^{tT} + \f12 it \sp{e^{t T}, \hat V}_+.
+  \approx \f12\sp{e^{tT}, e^{it\check V}}_+
+  \approx e^{tT} + \f12 it \sp{e^{t T}, \check V}_+.
 \end{align}

### 2019-04-17 soc_squeezing/collective_paper/collective_spin.tex
plot magnitude and phase of two-time correlator, rather than real/imaginary parts; fix commentary on two-time correlator plot (in collective spin paper) to reflect recent updates

@@
   \includegraphics{two_time.pdf}
-  \caption{Real (solid lines) and imaginary (dashed lines) parts of
-    the two-time connected correlator $C\p{t}$ defined in
+  \caption{Magnitude and phase of the two-time connected correlator
+    $C\p{t}\equiv\abs{C\p{t}}\exp\sp{i\phi\p{t}}$ defined in
     Eq.~\eqref{eq:two_time} for $N=10^4$ spins initially in the
@@
     evolution without decoherence (leftmost panel), as well as with
-    decoherence at rates $\gamma_\a$ or $\Gamma_\a$,
-    respectively corresponding to single-spin or collective jump
-    operators $s_\a^{(j)}$ or $S_\a$.  All axes in the sub-figures are
-    shared.  At short times, the correlations captured by $C\p{t}$ are
-    suppressed by single-spin decay ($\gamma_-$), collective decay
-    ($\Gamma_-$), and single-spin dephasing ($\gamma_\z$); these
-    correlations are relatively insensitive to collective dephasing
-    ($\Gamma_\z$), and can change sign or grow from spontaneous
-    single-spin ($\gamma_+$) or collective ($\Gamma_+$) excitation.}
+    decoherence at rates $\Gamma_\a$ or $\gamma_\a$, respectively
+    corresponding to collective or single-spin decoherence through
+    jump operators $S_\a$ or $s_\a^{(j)}$.  Collective decoherence
+    rates ($\Gamma_\a$) were chosen to be comparable to the OAT
+    squeezing strength $\chi$, while single-spin decoherence rates
+    ($\gamma_\a$) were scaled by an additional factor of $S=N/2$ in
+    order to yield comparable effects on the correlator $C\p{t}$ at
+    the times shown.  At short times, the correlations captured by
+    $C\p{t}$ are essentially insensitive to collective dephasing
+    ($\Gamma_\z$), while the magnitude $\abs{C\p{t}}$ is suppressed by
+    other forms of decoherence.  The type of decoherence (namely spin
+    decay, excitation, or dephasing) leaves a signature on the phase
+    $\phi\p{t}$ of $C\p{t}$.}
   \label{fig:two_time}
@@
 with and without various sources of decoherence.  The two-time
-correlator $C\p{t}$ degrades at short times under certain types of
-decoherence, but seems relatively insensitive to collective dephasing
-(with jump operator $\hat S_\z$), and can qualitatively change
-behavior or grow in the presence of single-spin or collective
-excitation (with jump operators $\hat s_+^{(j)}$ or $\hat S_+$).
+correlations captured by $C\p{t}\equiv\abs{C\p{t}}\exp\sp{i\phi\p{t}}$
+exhibit signatures of the type of decoherence present, with e.g.~clear
+dependence of the phase $\phi\p{t}$ on spin decay or excitation.
 Figure \ref{fig:two_time} serves as an example for the type of

### 2019-04-17 soc_squeezing/collective_paper/collective_spin.tex
changed order_cap in two-time correlator calculation to 20; renamed eq:TST in collective spin paper added a few missing hats on operators

@@
   \sum_{\v m, \v n} \hat\S_{\v m} T^k_{\v m\v n} \O_{\v n}\p{0}.
-  \label{eq:order_truncation}
-\end{align}
-We refer Eq.~\eqref{eq:order_truncation} as the truncated short-time
-(TST) expansion of Heisenberg operators.  Unlike the truncation in
+  \label{eq:TST}
+\end{align}
+We refer Eq.~\eqref{eq:TST} as the truncated short-time (TST)
+expansion of Heisenberg operators.  Unlike the truncation in
 Eq.~\eqref{eq:weight_truncation}, the nonzero matrix elements
-$T^k_{\v m\v n}$ for $k=0,1,\cdots,M$ in
-Eq.~\eqref{eq:order_truncation} tell us which operators are relevant
-for computing the expectation value $\bk{\hat\O\p{t}}$ to a fixed
-order $M$.  Note that using the relation
+$T^k_{\v m\v n}$ for $k=0,1,\cdots,M$ in Eq.~\eqref{eq:TST} tell us
+which operators are relevant for computing the expectation value
+$\bk{\hat\O\p{t}}$ to a fixed order $M$.  Note that using the relation
 $\hat\S_{\v m}^\dag=\hat\S_{\v m^*}$, which by Hermitian conjugation
@@
 $t$, there exists a truncation order $M_t$ for which the truncation
-error in Eq.~\eqref{eq:order_truncation} is negligibly small.  In
-practice, only a maximal truncation order $M_{\t{max}}$ is accessible
-with limited computational resources, such that the TST expansion in
-Eq.~\eqref{eq:order_truncation} only allows for computation of the
-expectation value $\bk{\hat\O\p{t}}$ to a maximal time
-$t_{M_{\t{max}}}$.  As we will see in the following section,
-$M_{\t{max}}=35$ will suffice for the computation of collective spin
-correlators up to times that are relevant e.g.~for spin squeezing
-protocols.
+error in Eq.~\eqref{eq:TST} is negligibly small.  In practice, only a
+maximal truncation order $M_{\t{max}}$ is accessible with limited
+computational resources, such that the TST expansion in
+Eq.~\eqref{eq:TST} only allows for computation of the expectation
+value $\bk{\hat\O\p{t}}$ to a maximal time $t_{M_{\t{max}}}$.  As we
+will see in the following section, $M_{\t{max}}=35$ will suffice for
+the computation of collective spin correlators up to times that are
+relevant e.g.~for spin squeezing protocols.
 
@@
 by one- and two-spin correlators of the form $\bk{\hat S_\a}$ and
-$\bk{\hat S_\a \hat S_\b}$.  For the unitary dynamics discussed in this
-work, these correlators are obtainable via exact simulations of
+$\bk{\hat S_\a \hat S_\b}$.  For the unitary dynamics discussed in
+this work, these correlators are obtainable via exact simulations of
 quantum dynamics in the $\p{N+1}$-dimensional Dicke manifold of states
@@
 this work, these exact and quantum trajectory simulations will be used
-to benchmark the TST expansion in Eq.~\eqref{eq:order_truncation}.
+to benchmark the TST expansion in Eq.~\eqref{eq:TST}.
 
@@
     using exact methods (solid lines), quantum trajectory simulations
-    (dots), and the TST expansion in Eq.~\eqref{eq:order_truncation}
-    with $M=35$ (darker dashed lines).  Solid circles mark the times
-    at which the TST expansion gives an unphysical result with
-    $\xi^2<0$.}
+    (dots), and the TST expansion in Eq.~\eqref{eq:TST} with $M=35$
+    (darker dashed lines).  Solid circles mark the times at which the
+    TST expansion gives an unphysical result with $\xi^2<0$.}
   \label{fig:benchmarking}
@@
 both benchmarking simulations and the TST expansion in
-Eq.~\eqref{eq:order_truncation} with $M=35$.  Squeezing is shown for
-both unitary dynamics (Figure \ref{fig:squeezing_coherent}), as well
-as non-unitary dynamics in the presence of spontaneous decay,
-excitation, and dephasing of individual spins at rates $\chi$ (Figure
+Eq.~\eqref{eq:TST} with $M=35$.  Squeezing is shown for both unitary
+dynamics (Figure \ref{fig:squeezing_coherent}), as well as non-unitary
+dynamics in the presence of spontaneous decay, excitation, and
+dephasing of individual spins at rates $\chi$ (Figure
 \ref{fig:squeezing_incoherent}), respectively described by the sets of
-jump operators $\A_\a=\set{\hat s_\a^{(j)}}$ with corresponding
+jump operators $\A_\a\equiv\set{\hat s_\a^{(j)}}$ with corresponding
 decoherence rates $\gamma_\a=\chi$ for $\a\in\set{-,+,\z}$.  The
@@
     at rates $\gamma_-=\gamma_+=\gamma_\z=100\chi$.  Computed using
-    the TST expansion in Eq.~\eqref{eq:order_truncation} with $M=35$.
-    Solid circles mark the times at which the TST expansion gives an
+    the TST expansion in Eq.~\eqref{eq:TST} with $M=35$.  Solid
+    circles mark the times at which the TST expansion gives an
     unphysical result with $\xi^2<0$.}
@@
     corresponding to collective or single-spin decoherence through
-    jump operators $S_\a$ or $s_\a^{(j)}$.  Collective decoherence
-    rates ($\Gamma_\a$) were chosen to be comparable to the OAT
-    squeezing strength $\chi$, while single-spin decoherence rates
-    ($\gamma_\a$) were scaled by an additional factor of $S=N/2$ in
-    order to yield comparable effects on the correlator $C\p{t}$ at
+    jump operators $\hat S_\a$ or $\hat s_\a^{(j)}$.  Collective
+    decoherence rates ($\Gamma_\a$) were chosen to be comparable to
+    the OAT squeezing strength $\chi$, while single-spin decoherence
+    rates ($\gamma_\a$) were scaled by an additional factor of $S=N/2$
+    in order to yield comparable effects on the correlator $C\p{t}$ at
     the times shown.  At short times, the correlations captured by
     $C\p{t}$ are essentially insensitive to collective dephasing
-    ($\Gamma_\z$), while the magnitude $\abs{C\p{t}}$ is suppressed by
-    other forms of decoherence.  The type of decoherence (namely spin
-    decay, excitation, or dephasing) leaves a signature on the phase
-    $\phi\p{t}$ of $C\p{t}$.}
+    ($\Gamma_\z$), while the magnitude $\abs{C\p{t}}$ is generally
+    suppressed by other forms of decoherence.  The general type of
+    decoherence (namely spin decay, excitation, or dephasing) also
+    leaves a signature on the phase $\phi\p{t}$ of $C\p{t}$.  Results
+    computed using the TST expansion in Eq.~\eqref{eq:TST} with
+    $M=20$.}
   \label{fig:two_time}

### 2019-04-18 soc_squeezing/collective_paper/collective_spin.tex
added citation to squeezing paper on arxiv; added tutorial for computing a correlator

@@
 \newcommand{\A}{\mathcal{A}}
-\newcommand{\C}{\mathcal{C}}
 \newcommand{\D}{\mathcal{D}}
@@
 
+\newcommand{\C}{\mathbb{C}}
 \newcommand{\N}{\mathbb{N}}
@@
 spin operator algebra; the calculation of these structure constants
-(in Appendices \ref{sec:identities}--\ref{sec:prod_general}) is one of
-the main technical results of this work, which we hope will empower
+(in Appendices \ref{sec:identities}--\ref{sec:general_product}) is one
+of the main technical results of this work, which we hope will empower
 both analytical and numerical studies of collective spin systems in
@@
 \end{align}
-with scalar coefficients $\O_{\v m}\in\mathbb{C}$.  If $\hat\O$ is
+with scalar coefficients $\O_{\v m}\in\C$.  If $\hat\O$ is
 self-adjoint, for example, then $\O_{\v m}^*=\O_{\v m^*}$ with
@@
 for a time derivative (super-)operator $\check T$ with matrix elements
-$T_{\v m\v n}\in\mathbb{C}$ defined by
+$T_{\v m\v n}\in\C$ defined by
 \begin{align}
@@
 would be described by the set of jump operators
-$\A_-=\set{\hat s_-^{(j)}:j=1,2,\cdots,N}$.  The commutator in
+$\A_-\equiv\set{\hat s_-^{(j)}:j=1,2,\cdots,N}$.  The commutator in
 Eq.~\eqref{eq:time_deriv} can be computed by expanding the product
@@
 $f_{\v\ell\v m\v n}\in\mathbb{R}$ that we work out in Appendices
-\ref{sec:identities}--\ref{sec:prod_general}, and the effects of
+\ref{sec:identities}--\ref{sec:general_product}, and the effects of
 decoherence from jump operators (i.e.~elements of $\A$) of the form
@@
 where the matrix elements $T^k_{\v m\v n}$ of the $k$-th time
-derivative operator are
+derivative operator $T^k$ are
 \begin{align}
@@
 the computation of collective spin correlators up to times that are
-relevant e.g.~for spin squeezing protocols.
+relevant e.g.~for spin squeezing protocols.  Appendix
+\ref{sec:tutorial} provides a pedagogical tutorial for computing
+correlators using the TST expansion.
 
@@
     ($\Gamma_\z$), while the magnitude $\abs{C\p{t}}$ is generally
-    suppressed by other forms of decoherence.  The general type of
-    decoherence (namely spin decay, excitation, or dephasing) also
-    leaves a signature on the phase $\phi\p{t}$ of $C\p{t}$.  Results
-    computed using the TST expansion in Eq.~\eqref{eq:TST} with
-    $M=20$.}
+    suppressed by other forms of decoherence.  Whether decoherence
+    occurs through spin decay, excitation, or dephasing also leaves a
+    signature on the phase $\phi\p{t}$ of $C\p{t}$.  Results computed
+    using the TST expansion in Eq.~\eqref{eq:TST} with $M=20$.}
   \label{fig:two_time}
@@
 \section{Product of arbitrary ordered collective spin operators}
-\label{sec:prod_general}
+\label{sec:general_product}
 
@@
 
+\section{Computing correlators with the truncated short-time (TST)
+  expansion}
+\label{sec:tutorial}
+
+Here, we provide a pedagogical tutorial for computing correlators
+using the truncated short-time TST expansion.  For concreteness, we
+nominally consider $N$ spins evolving under the one-axis twisting
+(OAT) Hamiltonian
+\begin{align}
+  H_{\t{OAT}} = \chi S_\z^2,
+\end{align}
+additionally subject to spontaneous single-spin decay at rate
+$\gamma_-$, with jump operators $\A_-=\set{s_-^{(j)}:j=1,2,\cdots,N}$.
+The equation of motion for a Heisenberg operator
+$S_+^\ell S_\z^m S_-^n$ is
+\begin{align}
+  \f{d}{dt} \p{S_+^\ell S_\z^m S_-^n}
+  = i\chi\sp{S_\z^2, S_+^\ell S_\z^m S_-^n}_-
+  + \gamma_- \D\p{\A_-} \p{S_+^\ell S_\z^m S_-^n},
+  \label{eq:tutorial_EOM}
+\end{align}
+where we can use the results in appendices \ref{sec:general_product}
+and \ref{sec:decoherence_single} respectively to evaluate the
+commutator $\sp{S_\z^2, S_+^\ell S_\z^m S_-^n}_-$ and dissipator
+$\D\p{\A_-} \p{S_+^\ell S_\z^m S_-^n}$ in \eqref{eq:tutorial_EOM} (see
+\eqref{eq:general_product} and \eqref{eq:decay_diff} in particular),
+giving us
+\begin{align}
+  \f{d}{dt} \p{S_+^\ell S_\z^m S_-^n} =
+  i\chi \p{\ell-n} S_+^\ell \p{\ell+n+2S_\z} S_\z^m S_-^n
+  + \gamma_- S_+^\ell \sp{\p{S+S_\z}\p{-1+S_\z}^m
+    - \p{S+\f{\ell+n}{2}+S_\z} S_\z^m} S_-^n.
+  \label{eq:tutorial_EOM_expanded}
+\end{align}
+In practice, we do not want to keep track of such an expansion by
+hand, especially in the case of e.g.~the two-axis twisting (TAT) and
+twist-and-turn (TNT) models with more general types of decoherence,
+for which the analogue of \eqref{eq:tutorial_EOM_expanded} may take
+several lines just to write out in full.  Defining the operators
+$\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$ with
+$\v m\equiv\p{m_+,m_\z,m_-}$ for shorthand, we note that the vector
+space spanned by $\set{\S_{\v m}}$ is closed under time evolution.  We
+therefore expand
+\begin{align}
+  \f{d}{dt} \S_{\v n} \equiv T \S_{\v n}
+  = \sum_{\v m} \S_{\v m} T_{\v m\v n},
+  \label{eq:tutorial_EOM_general}
+\end{align}
+where $T$ is a super-operator that generates time evolution for
+Heisenberg operators.  In the present example, the matrix elements
+$T_{\v m\v n}\in\C$ of $T$ are defined by
+\eqref{eq:tutorial_EOM_expanded} and \eqref{eq:tutorial_EOM_general}.
+For any Hamiltonian $H$ with decoherence characterized by sets of jump
+operators $\A$ and decoherence rates $\gamma_A$, the matrix elements
+$T_{\v m\v n}$ are more generally defined by
+\begin{align}
+  T \S_{\v n}
+  = i\sp{H,\S_{\v n}}_- + \sum_\A \gamma_\A \D\p{\A} \S_{\v n}
+  = \sum_{\v m} \S_{\v m} T_{\v m\v n}.
+\end{align}
+The results in Appendices \ref{sec:general_product},
+\ref{sec:decoherence_single}, and \ref{sec:decoherence_collective} can
+be used to write model-agnostic codes that compute matrix elements
+$T_{\v m\v n}$, taking a particular Hamiltonian $H$ and decoherence
+processes $\set{\p{\A,\gamma_\A}}$ as inputs.
+
+In order to compute a quantity such as spin squeezing, we need to
+compute correlators of the form $\bk{\S_{\v n}\p{t}}$, where
+$\S_{\v n}\p{t}$ is the time-evolved Heisenberg operator corresponding
+to $\S_{\v n}$, with $\S_{\v n}\p{0}\equiv\S_{\v n}$.  The order-$M$
+truncated short-time (TST) expansion takes
+\begin{align}
+  \bk{\S_{\v n}\p{t}}
+  \to \sum_{k=0}^M \f{t^k}{k!} \sum_{\v m} \bk{\S_{\v m}} T^k_{\v m\v n},
+  \label{eq:tutorial_TST}
+\end{align}
+where $T^k_{\v m\v n}$ are matrix elements of the $k$-th time
+derivative operator, given by
+\begin{align}
+  T^0_{\v m\v n} \equiv
+    \begin{cases}
+      1 & \v m = \v n, \\
+      0 & \t{otherwise}
+    \end{cases},
+  &&
+  T^1_{\v m\v n} \equiv T_{\v m\v n},
+  &&
+  T^{k>1}_{\v m\v n}
+  \equiv \sum_{\v p_1,\v p_2,\cdots,\v p_{k-1}}
+  T_{\v m\v p_{k-1}} \cdots T_{\v p_3\v p_2}
+  T_{\v p_2\v p_1} T_{\v p_1\v n}.
+\end{align}
+Matrix elements $T^k_{\v m\v n}$ and initial-time expectation values
+$\bk{\S_{\v m}}$ are thus computed on an as-needed basis for any
+particular correlator $\bk{\S_{\v n}\p{t}}$ of interest, and combined
+according to \eqref{eq:tutorial_TST}.  Note that initial-time
+expectation values $\bk{\S_{\v m}}$ are an input to the TST expansion,
+and need to be computed separately for any particular initial state of
+interest; expectation values with respect to spin-polarized (Gaussian)
+states are provided in Appendix \ref{sec:initial_conditions}.  In
+practice, we further collect terms in \eqref{eq:tutorial_TST} to write
+\begin{align}
+  \bk{\S_{\v n}\p{t}} \to \sum_{k=0}^M c_{\v n k} t^k,
+  &&
+  c_{\v n k} \equiv \f1{k!} \sum_{\v m} \bk{\S_{\v m}} T^k_{\v m\v n},
+\end{align}
+where $c_{\v n k}$ are time-independent coefficients for the
+polynomial time expansion of $\bk{\S_{\v n}\p{t}}$.  After computing
+the coefficients $c_{\v n k}$, there is only negligible computational
+overhead to compute the correlator $\bk{\S_{\v n}\p{t}}$ for any time
+$t$ up to the breakdown time $t_M$ of an order-$M$ TST expansion.
+
 \section{Short-time linear response and two-time correlators}
@@
 Here, we discuss the appearance of two-time correlation functions in
-the short-time linear response of Heisenberg operators to
-perturbations of a Hamiltonian.  Consider an initial Hamiltonian $H$
-perturbed by an operator $V$ with $\norm{V}\ll\norm{H}$, where
-$\norm{\O}$ denotes the operator norm of $\O$, such that the net
-Hamiltonian is $\widetilde H=H+V$.  We denote the generator of
-Heisenberg time evolution under the perturbed (unperturbed)
-Hamiltonian by $\widetilde T$ ($T$).  These generators are related by
+the short-time linear response of correlators to perturbations of a
+Hamiltonian.  Consider an initial Hamiltonian $H$ perturbed by an
+operator $V$ with $\norm{V}\ll\norm{H}$, where $\norm{\O}$ denotes the
+operator norm of $\O$, such that the net Hamiltonian is
+$\widetilde H=H+V$.  We denote the generator of Heisenberg time
+evolution under the perturbed (unperturbed) Hamiltonian by
+$\widetilde T$ ($T$).  These generators are related by
 \begin{align}

### 2019-04-20 soc_squeezing/collective_paper/collective_spin.tex
refactored code in two-time correlator notebook
minor wording change in an appendix

@@
     signature on the phase $\phi\p{t}$ of $C\p{t}$.  Results computed
-    using the TST expansion in Eq.~\eqref{eq:TST} with $M=20$.}
+    using the TST expansion in Eq.~\eqref{eq:TST} with $M=35$.}
   \label{fig:two_time}
@@
 expectation values $\bk{\S_{\v m}}$ are an input to the TST expansion,
-and need to be computed separately for any particular initial state of
-interest; expectation values with respect to spin-polarized (Gaussian)
-states are provided in Appendix \ref{sec:initial_conditions}.  In
-practice, we further collect terms in \eqref{eq:tutorial_TST} to write
+and need to be computed separately for any given initial state;
+expectation values with respect to spin-polarized (Gaussian) states
+are provided in Appendix \ref{sec:initial_conditions}.  In practice,
+we further collect terms in \eqref{eq:tutorial_TST} to write
 \begin{align}

### 2019-04-20 soc_squeezing/collective_paper/collective_spin.tex
added OTOC figure and commentary

@@
   We present a new method to compute short-time expectation values in
-  large ensembles of collective spin systems with generic Markovian
-  decoherence.  Our method is based on a Taylor expansion of a formal
-  solution to the equations of motion for Heisenberg operators.  This
-  expansion can be truncated at finite order to obtain virtually exact
-  results at short times that are relevant for metrological
-  applications such as spin squeezing.  In order to evaluate the
-  expansion for Heisenberg operators, we compute the relevant
-  structure constants of a collective operator algebra.  We
-  demonstrate the utility of our method by computing spin squeezing
-  and two-time correlation functions in strong-decoherence regimes
-  that are inaccessible to existing numerical methods.  Our method can
-  be straightforwardly generalized to the case of a collective spin
-  coupled to bosonic modes, relevant for trapped ion and cavity QED
-  experiments, and may be used to compute quantities such as
-  out-of-time-ordered correlators that can feature short-time
-  signatures of quantum chaos and information scrambling.
+  large collective spin systems with generic Markovian decoherence.
+  Our method is based on a Taylor expansion of a formal solution to
+  the equations of motion for Heisenberg operators.  This expansion
+  can be truncated at finite order to obtain virtually exact results
+  at short times that are relevant for metrological applications such
+  as spin squeezing.  In order to evaluate the expansion for
+  Heisenberg operators, we compute the relevant structure constants of
+  a collective operator algebra.  We demonstrate the utility of our
+  method by computing spin squeezing, two-time correlation functions,
+  and out-of-time-ordered correlators for $10^4$ spins in
+  strong-decoherence regimes that are otherwise inaccessible via
+  existing numerical methods.  Our method can be straightforwardly
+  generalized to the case of a collective spin coupled to bosonic
+  modes, relevant for trapped ion and cavity QED experiments, and may
+  be used to investigate short-time signatures of quantum chaos and
+  information scrambling.
 \end{abstract}
@@
 decoherence in Ref.~[\citenum{he2019engineering}], exemplifying a
-concrete application of the TST expansion and the collective-spin
-structure constants calculated in this work.
+concrete and practical application of the TST expansion and the
+collective-spin structure constants calculated in this work.
+
+\section{Two-time correlation functions and out-of-time-ordered
+  correlators}
+\label{sec:multi_time}
 
@@
   \centering
-  \includegraphics{two_time.pdf}
-  \caption{Magnitude and phase of the two-time connected correlator
+  \subfloat[Two-time correlator
   $C\p{t}\equiv\abs{C\p{t}}\exp\sp{i\phi\p{t}}$ defined in
-    Eq.~\eqref{eq:two_time} for $N=10^4$ spins initially in the
-    polarized state $\ket{\X}\propto\p{\ket\up+\ket\dn}^{\otimes N}$
-    evolving under the squeezing Hamiltonians in
-    Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT}.  Results are shown for time
-    evolution without decoherence (leftmost panel), as well as with
-    decoherence at rates $\Gamma_\a$ or $\gamma_\a$, respectively
-    corresponding to collective or single-spin decoherence through
-    jump operators $\hat S_\a$ or $\hat s_\a^{(j)}$.  Collective
-    decoherence rates ($\Gamma_\a$) were chosen to be comparable to
-    the OAT squeezing strength $\chi$, while single-spin decoherence
-    rates ($\gamma_\a$) were scaled by an additional factor of $S=N/2$
-    in order to yield comparable effects on the correlator $C\p{t}$ at
-    the times shown.  At short times, the correlations captured by
-    $C\p{t}$ are essentially insensitive to collective dephasing
-    ($\Gamma_\z$), while the magnitude $\abs{C\p{t}}$ is generally
+  Eq.~\eqref{eq:two_time}. \label{fig:two_time}]
+  {\includegraphics{two_time.pdf}} \\
+  \subfloat[Squared commutator $D\p{t}$ defined in
+  Eq.~\eqref{eq:OTOC}. \label{fig:OTOC}]
+  {\includegraphics{OTOC.pdf}}
+  \caption{The two-time connected correlator $C\p{t}$ and squared
+    commutator $D\p{t}$, respectively defined in
+    Eqs.~\eqref{eq:two_time} and \eqref{eq:OTOC}, for $N=10^4$ spins
+    initially in the polarized state
+    $\ket{\X}\propto\p{\ket\up+\ket\dn}^{\otimes N}$ evolving under
+    the squeezing Hamiltonians in Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT}.
+    Results are shown for time evolution without decoherence (leftmost
+    panels), as well as with decoherence at rates $\Gamma_\a$ or
+    $\gamma_\a$, respectively corresponding to collective or
+    single-spin decoherence through jump operators $\hat S_\a$ or
+    $\set{\hat s_\a^{(j)}}$.  Collective decoherence rates
+    ($\Gamma_\a$) were chosen to be comparable to the OAT squeezing
+    strength $\chi$, while single-spin decoherence rates ($\gamma_\a$)
+    were scaled by an additional factor of $S=N/2$ in order to yield
+    comparable effects on the correlators $C\p{t}$ and $D\p{t}$ at the
+    times shown.  At short times, the multi-time correlations captured
+    by $C\p{t}$ and $D\p{t}$ are essentially insensitive to collective
+    dephasing ($\Gamma_\z$), while their magnitudes are generally
     suppressed by other forms of decoherence.  Whether decoherence
@@
     signature on the phase $\phi\p{t}$ of $C\p{t}$.  Results computed
-    using the TST expansion in Eq.~\eqref{eq:TST} with $M=35$.}
-  \label{fig:two_time}
+    using the TST expansion in Eq.~\eqref{eq:TST} with $M=20$.}
+  \label{fig:multi_time}
 \end{figure*}
-
-\section{Two-time correlation functions}
-\label{sec:two_time}
 
@@
 accessible via the TST expansion of Heisenberg operators, we consider
-the calculation of collective-spin two-time correlation functions.  In
-particular, we consider the effect of decoherence on short-time
-behavior of the two-time connected correlator
+the calculation of two-time correlation functions and
+out-of-time-ordered correlators (OTOCs).  In particular, we consider
+the effect of decoherence on short-time behavior of the two-time
+connected correlator
 \begin{align}
   C\p{t}
-  \equiv \f1S\sp{\bk{\hat S_+\p{t} \hat S_-\p{0}}
-    - \bk{\hat S_+\p{t}}\bk{\hat S_-\p{0}}}
+  \equiv \f1S\p{\bk{\hat S_+\p{t} \hat S_-\p{0}}
+    - \bk{\hat S_+\p{t}}\bk{\hat S_-\p{0}}},
   \label{eq:two_time}
 \end{align}
+and the expectation value of a squared commutator,
+\begin{align}
+  D\p{t}
+  \equiv \f1{S^2}\Braket{\sp{\hat S_+\p{t},\hat S_-\p{0}}_-^\dag
+    \sp{\hat S_+\p{t},\hat S_-\p{0}}_-},
+  \label{eq:OTOC}
+\end{align}
 in the context of the squeezing models in Section \ref{sec:squeezing}.
-In an equilibrium setting, correlation functions of this sort contain
-information about the linear response of Heisenberg operators to
-perturbations of a system; in a non-equilibrium setting, they
-contribute to short-time linear response (see Appendix
-\ref{sec:linear_response}).  Similar correlators have made appearances
-as order parameters that diagnose time-crystalline phases of
-matter\cite{tucker2018shattered}, and related multi-time correlators,
-known as out-of-time-ordered correlators, are commonly examined for
-signatures of quantum chaos and information
-scrambling\cite{maldacena2016bound}.
-
-Figure \ref{fig:two_time} shows the behavior of $C\p{t}$ for $N=10^4$
-spins, initially in the state
+In an equilibrium setting, correlation functions similar to that in
+Eq.~\eqref{eq:two_time} contain information about the linear response
+of Heisenberg operators to perturbations of a system; in a
+non-equilibrium setting, they contribute to short-time linear response
+(see Appendix \ref{sec:linear_response}).  Similar correlators have
+made appearances as order parameters that diagnose time-crystalline
+phases of matter\cite{tucker2018shattered}.  OTOCs such as that in
+Eq.~\eqref{eq:OTOC}, meanwhile, are commonly examined for signatures
+of quantum chaos and information scrambling\cite{maldacena2016bound,
+  swingle2018unscrambling, garcia-mata2018chaos}.
+
+Figure \ref{fig:multi_time} shows the behavior of $C\p{t}$ and
+$D\p{t}$ for $N=10^4$ spins, initially in the state
 $\ket\X\propto\p{\ket\up+\ket\dn}^{\otimes N}$, evolving under the
 squeezing Hamiltonians in Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT} both
-with and without various sources of decoherence.  The two-time
+with and without various sources of decoherence.  At short times, both
+correlators are insensitive to collective dephasing, while their
+magnitudes are suppressed by other types of decoherence.  The
 correlations captured by $C\p{t}\equiv\abs{C\p{t}}\exp\sp{i\phi\p{t}}$
-exhibit signatures of the type of decoherence present, with e.g.~clear
-dependence of the phase $\phi\p{t}$ on spin decay or excitation.
-Figure \ref{fig:two_time} serves as an example for the type of
-behavior that is accessible at short times with the TST expansion.
-This example is straightforward to extend to multi-time and
-out-of-time-ordered correlators, as well as equilibrium settings.
+also exhibit signatures of the type of decoherence present, with
+e.g.~clear dependence of the phase $\phi\p{t}$ on spin decay or
+excitation.  Figure \ref{fig:multi_time} serves as an example for the
+type of behavior that is accessible at short times with the TST
+expansion.  These examples are straightforward extend to equilibrium
+settings and spin-boson systems.
 
@@
 
-We have presented a new method for computing correlators in collective
-spin systems at short times.  This method is based on truncating a
-short-time expansion of Heisenberg operators, and can access
+We have presented a new method for computing correlators at short
+times in collective spin systems.  This method is based on truncating
+a short-time expansion of Heisenberg operators, and can access
 correlators on time scales that are relevant to metrological
@@
 may be used to benchmark the short-time effects of decoherence, or
-compute quantities such as out-of-time-ordered correlators that
-diagnose the onset of quantum chaos and information scrambling.
+study the onset of quantum chaos and information scrambling.
 

### 2019-04-23 soc_squeezing/collective_paper/collective_spin.tex
minor corrections and wording changes

@@
   \to \sum_{w\p{\v m}<W} \hat\S_{\v m}
-  \sum_{\v n} T_{\v m\v n} \hat\O_{\v n}\p{t}
+  \sum_{\v n} T_{\v m\v n} \O_{\v n}\p{t}
   \label{eq:weight_truncation}
@@
   \subfloat[Squared commutator $D\p{t}$ defined in
-  Eq.~\eqref{eq:OTOC}. \label{fig:OTOC}]
+  Eq.~\eqref{eq:four_point}. \label{fig:four_point}]
   {\includegraphics{OTOC.pdf}}
@@
     commutator $D\p{t}$, respectively defined in
-    Eqs.~\eqref{eq:two_time} and \eqref{eq:OTOC}, for $N=10^4$ spins
-    initially in the polarized state
+    Eqs.~\eqref{eq:two_time} and \eqref{eq:four_point}, for $N=10^4$
+    spins initially in the polarized state
     $\ket{\X}\propto\p{\ket\up+\ket\dn}^{\otimes N}$ evolving under
@@
 
+% TODO: in contrast to typical OTOCs, here the operators "almost commute" -- up to 1/N terms
+
+% TODO: mention short-time expansion, see only polynomial growth
+
+% TODO: figures 3a, 3b: 3 panels each; no decoherence, decoherence in fig. 2, and strong decoherence
+
+% TODO: add vertical axis labels for G_- and g_- (?)
+
 As a final example of collective-spin physics that is numerically
@@
     \sp{\hat S_+\p{t},\hat S_-\p{0}}_-},
-  \label{eq:OTOC}
+  \label{eq:four_point}
 \end{align}
@@
 made appearances as order parameters that diagnose time-crystalline
-phases of matter\cite{tucker2018shattered}.  OTOCs such as that in
-Eq.~\eqref{eq:OTOC}, meanwhile, are commonly examined for signatures
-of quantum chaos and information scrambling\cite{maldacena2016bound,
-  swingle2018unscrambling, garcia-mata2018chaos}.
+phases of matter\cite{tucker2018shattered}.  Squared commutators such
+as that in Eq.~\eqref{eq:four_point}, meanwhile, are commonly examined
+for signatures of quantum chaos and information
+scrambling\cite{maldacena2016bound, swingle2018unscrambling,
+  garcia-mata2018chaos}.
 
@@
   \bk{\S_{\v n}\p{t}}
-  \to \sum_{k=0}^M \f{t^k}{k!} \sum_{\v m} \bk{\S_{\v m}} T^k_{\v m\v n},
+  \to \sum_{k=0}^M \f{t^k}{k!}
+  \sum_{\v m} \bk{\S_{\v m}\p{0}} T^k_{\v m\v n},
   \label{eq:tutorial_TST}
@@
 Matrix elements $T^k_{\v m\v n}$ and initial-time expectation values
-$\bk{\S_{\v m}}$ are thus computed on an as-needed basis for any
-particular correlator $\bk{\S_{\v n}\p{t}}$ of interest, and combined
-according to \eqref{eq:tutorial_TST}.  Note that initial-time
-expectation values $\bk{\S_{\v m}}$ are an input to the TST expansion,
-and need to be computed separately for any given initial state;
-expectation values with respect to spin-polarized (Gaussian) states
-are provided in Appendix \ref{sec:initial_conditions}.  In practice,
-we further collect terms in \eqref{eq:tutorial_TST} to write
+$\bk{\S_{\v m}\p{0}}$ are thus computed as needed for any particular
+correlator $\bk{\S_{\v n}\p{t}}$ of interest, and combined according
+to \eqref{eq:tutorial_TST}.  Note that initial-time expectation values
+$\bk{\S_{\v m}\p{0}}$ are an input to the TST expansion, and need to
+be computed separately for any given initial state; expectation values
+with respect to spin-polarized (Gaussian) states are provided in
+Appendix \ref{sec:initial_conditions}.  In practice, we further
+collect terms in \eqref{eq:tutorial_TST} to write
 \begin{align}
@@
   &&
-  c_{\v n k} \equiv \f1{k!} \sum_{\v m} \bk{\S_{\v m}} T^k_{\v m\v n},
+  c_{\v n k}
+  \equiv \f1{k!} \sum_{\v m} \bk{\S_{\v m}\p{0}} T^k_{\v m\v n},
 \end{align}

### 2019-04-28 soc_squeezing/collective_paper/collective_spin.tex
changes to figures and associated commentary in collective spin paper

@@
 
-\title{Short-time simulations of open collective quantum spin systems}
+\title{Short-time expansion of Heisenberg operators in open collective
+  quantum spin systems}
 
@@
   = \f{N}{\lvert\braket{\uv S}\rvert^2}
-  \min_{\uv n\perp\bk{\uv S}} \Braket{\p{\uv S\cdot\uv n}^2},
+  \min_{\substack{\v v\perp\bk{\uv S}\\\v v\cdot\v v=1}}
+    \Braket{\p{\uv S\cdot\v v}^2},
   \label{eq:squeezing}
@@
 spin operator-valued vector, and the minimization is performed over
-all unit vectors $\uv n$ in the plane orthogonal to the mean spin
-vector $\bk{\uv S}$.  This squeezing parameter is entirely determined
-by one- and two-spin correlators of the form $\bk{\hat S_\a}$ and
+all unit vectors $\v v$ orthogonal to the mean spin vector
+$\bk{\uv S}$.  This squeezing parameter is entirely determined by one-
+and two-spin correlators of the form $\bk{\hat S_\a}$ and
 $\bk{\hat S_\a \hat S_\b}$.  For the unitary dynamics discussed in
@@
   \subfloat[Squeezing with unitary dynamics
-  \label{fig:squeezing_coherent}]
-  {\includegraphics{coherent.pdf}} \\
+  \label{fig:squeezing_unitary}]
+  {\includegraphics{unitary.pdf}} \\
   \subfloat[Squeezing with decoherence:
@@
   \caption{Spin squeezing of $N=10^4$ spins initially in $\ket\X$
-    under \sref{fig:squeezing_coherent} unitary and
+    under \sref{fig:squeezing_unitary} unitary and
     \sref{fig:squeezing_incoherent} non-unitary dynamics, computed
@@
 Eq.~\eqref{eq:TST} with $M=35$.  Squeezing is shown for both unitary
-dynamics (Figure \ref{fig:squeezing_coherent}), as well as non-unitary
+dynamics (Figure \ref{fig:squeezing_unitary}), as well as non-unitary
 dynamics in the presence of spontaneous decay, excitation, and
@@
 expansion gives essentially exact results right up until a sudden and
-drastic departure that can be diagnosed by inspection.  This drastic
-departure occurs because the TST expansion neglects high-weight
-(i.e.~large-$\abs{\v m}$) operators $\hat\S_{\v m}$ whose
+drastic departure that can be diagnosed by inspection.
+% todo: comment on diagnosing breakdown by examining individual correlators, rather than squeezing
+This drastic departure occurs because the TST expansion neglects
+high-weight (i.e.~large-$\abs{\v m}$) operators $\hat\S_{\v m}$ whose
 contributions to a Heisenberg operator $\hat\O\p{t}$ of interest
@@
 $\sim10^4$ CPU hours to compute on standard modern computing hardware;
-the bulk of this time is spent performing sparse matrix
+the bulk of this time is spent performing sparse matrix-vector
 multiplication, leaving little room to further optimize run time.
@@
 
-\begin{figure*}
-  \centering
-  \subfloat[Two-time correlator
-  $C\p{t}\equiv\abs{C\p{t}}\exp\sp{i\phi\p{t}}$ defined in
-  Eq.~\eqref{eq:two_time}. \label{fig:two_time}]
-  {\includegraphics{two_time.pdf}} \\
-  \subfloat[Squared commutator $D\p{t}$ defined in
-  Eq.~\eqref{eq:four_point}. \label{fig:four_point}]
-  {\includegraphics{OTOC.pdf}}
-  \caption{The two-time connected correlator $C\p{t}$ and squared
-    commutator $D\p{t}$, respectively defined in
-    Eqs.~\eqref{eq:two_time} and \eqref{eq:four_point}, for $N=10^4$
-    spins initially in the polarized state
-    $\ket{\X}\propto\p{\ket\up+\ket\dn}^{\otimes N}$ evolving under
-    the squeezing Hamiltonians in Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT}.
-    Results are shown for time evolution without decoherence (leftmost
-    panels), as well as with decoherence at rates $\Gamma_\a$ or
-    $\gamma_\a$, respectively corresponding to collective or
-    single-spin decoherence through jump operators $\hat S_\a$ or
-    $\set{\hat s_\a^{(j)}}$.  Collective decoherence rates
-    ($\Gamma_\a$) were chosen to be comparable to the OAT squeezing
-    strength $\chi$, while single-spin decoherence rates ($\gamma_\a$)
-    were scaled by an additional factor of $S=N/2$ in order to yield
-    comparable effects on the correlators $C\p{t}$ and $D\p{t}$ at the
-    times shown.  At short times, the multi-time correlations captured
-    by $C\p{t}$ and $D\p{t}$ are essentially insensitive to collective
-    dephasing ($\Gamma_\z$), while their magnitudes are generally
-    suppressed by other forms of decoherence.  Whether decoherence
-    occurs through spin decay, excitation, or dephasing also leaves a
-    signature on the phase $\phi\p{t}$ of $C\p{t}$.  Results computed
-    using the TST expansion in Eq.~\eqref{eq:TST} with $M=20$.}
-  \label{fig:multi_time}
-\end{figure*}
-
-% TODO: in contrast to typical OTOCs, here the operators "almost commute" -- up to 1/N terms
-
-% TODO: mention short-time expansion, see only polynomial growth
-
-% TODO: figures 3a, 3b: 3 panels each; no decoherence, decoherence in fig. 2, and strong decoherence
-
-% TODO: add vertical axis labels for G_- and g_- (?)
-
 As a final example of collective-spin physics that is numerically
@@
 scrambling\cite{maldacena2016bound, swingle2018unscrambling,
-  garcia-mata2018chaos}.
+  garcia-mata2018chaos}.  In typical scenarios, such squared
+commutators initially vanish by construction due to a choice of
+operators with spatially separated non-trivial support.  Collective
+spin systems, however, have no intrinsic notion of locality or spatial
+separation.  In our case, therefore, with the choice of initial state
+$\ket\X\propto\p{\ket\up+\ket\dn}^{\otimes N}$ we merely have
+$D\p{0}\sim1/N$.
+
+\begin{figure}
+  \centering
+  \subfloat[Two-time correlator
+  $C\p{t}\equiv\abs{C\p{t}}\exp\sp{i\phi\p{t}}$. \label{fig:two_time}]
+  {\includegraphics{two_time.pdf}} \\
+  \subfloat[Squared commutator $D\p{t}$. \label{fig:four_point}]
+  {\includegraphics{FPC.pdf}}
+  \caption{The two-time connected correlator $C\p{t}$ and squared
+    commutator $D\p{t}$, respectively defined in
+    Eqs.~\eqref{eq:two_time} and \eqref{eq:four_point}, for $N=10^4$
+    spins initially in the polarized state
+    $\ket{\X}\propto\p{\ket\up+\ket\dn}^{\otimes N}$ evolving under
+    the squeezing Hamiltonians in Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT}.
+    Results are shown for both unitary dynamics (solid lines) and
+    non-unitary dynamics with $\gamma_-=\gamma_+=\gamma_\z=100\chi$
+    (dashed lines), computed using the TST expansion in
+    Eq.~\eqref{eq:TST} with $M=20$.}
+  \label{fig:multi_time}
+\end{figure}
+
+% todo: add commentary/discussion about growth of C(t)
 
 Figure \ref{fig:multi_time} shows the behavior of $C\p{t}$ and
-$D\p{t}$ for $N=10^4$ spins, initially in the state
-$\ket\X\propto\p{\ket\up+\ket\dn}^{\otimes N}$, evolving under the
-squeezing Hamiltonians in Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT} both
-with and without various sources of decoherence.  At short times, both
-correlators are insensitive to collective dephasing, while their
-magnitudes are suppressed by other types of decoherence.  The
-correlations captured by $C\p{t}\equiv\abs{C\p{t}}\exp\sp{i\phi\p{t}}$
-also exhibit signatures of the type of decoherence present, with
-e.g.~clear dependence of the phase $\phi\p{t}$ on spin decay or
-excitation.  Figure \ref{fig:multi_time} serves as an example for the
-type of behavior that is accessible at short times with the TST
-expansion.  These examples are straightforward extend to equilibrium
-settings and spin-boson systems.
+$D\p{t}$ for $N=10^4$ spins, initially in the state $\ket\X$, evolving
+under the squeezing Hamiltonians in
+Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT} both with and without spontaneous
+decay, excitation, and dephasing of individual spins at rates
+$\gamma_-=\gamma_+=\gamma_\z=100\chi$.  In the case of TAT and TNT,
+after an initial ramp-up period these correlators grow exponentially
+as $\sim\exp\p{a\chi t}$ with $a\sim N$.  The growth of these
+correlators is definitively sub-exponential for OAT.  In the case of
+unitary time evolution under OAT in particular, to an excellent
+approximation we find that $D\p{t}=D\p{0}+\p{b\chi t}^2$ with
+$b=N+O\p{1}$.  As may be expected, the growth rates of $C\p{t}$ and
+$D\p{t}$ are generally suppressed by decoherence.  Figure
+\ref{fig:multi_time} serves as an example for the type of behavior
+that is accessible at short times with the TST expansion.  These
+examples are straightforward extend to equilibrium settings and
+spin-boson systems.
 
@@
 For any Hamiltonian $H$ with decoherence characterized by sets of jump
-operators $\A$ and decoherence rates $\gamma_A$, the matrix elements
+operators $\A$ and decoherence rates $\gamma_\A$, the matrix elements
 $T_{\v m\v n}$ are more generally defined by

### 2019-04-28 soc_squeezing/scripts/plot_squeezing.py

@@ -19,7 +19,7 @@ save = "save" in sys.argv
 data_dir = "../data/squeezing/"
 fig_dir = "../figures/squeezing/"
 
-figsize = (3,2.4)
+figsize = (3,2.2)
 params = { "text.usetex" : True,
            "font.size" : 8 }
 plt.rcParams.update(params)
@@ -250,7 +250,7 @@ def darken_color(color, darken_val = 0.08):
                            for jj in (1, 3 ,5) ])
     return tuple( min(max(cc-darken_val,0), 1) for cc in color )
 
-### coherent evolution
+### unitary evolution
 
 plt.figure(figsize = figsize)
 color = {}
@@ -267,15 +267,15 @@ for method in methods:
                      [time_sqz_C_trunc[method][1][positive_vals-1]],
                      "o", color = darken_color(colors[method]))
 
-plt.xlabel(r"$N\chi t$")
-plt.ylabel(r"$\xi^2$")
+plt.xlabel(r"Time ($N\chi t$)")
+plt.ylabel(r"Squeezing ($\xi^2$)")
 
 plt.xlim(0, max_plot_time)
 plt.ylim(*ylims(time_sqz_C_exact,time_sqz_C_trunc))
 
-plt.legend(loc = "best")
+plt.legend(loc = "best", bbox_to_anchor = (0.5, 0.5, 0.5, 0.5))
 plt.tight_layout()
-if save: plt.savefig(fig_dir + "coherent.pdf")
+if save: plt.savefig(fig_dir + "unitary.pdf")
 
 
 ### evolution with weak decoherence
@@ -295,8 +295,8 @@ for method in methods:
                      [time_sqz_D_trunc[method][1][positive_vals-1]],
                      "o", color = darken_color(colors[method]))
 
-plt.xlabel(r"$N\chi t$")
-plt.ylabel(r"$\xi^2$")
+plt.xlabel(r"Time ($N\chi t$)")
+plt.ylabel(r"Squeezing ($\xi^2$)")
 
 plt.xlim(0, max_plot_time)
 plt.ylim(*ylims(time_sqz_D_exact,time_sqz_D_trunc))
@@ -319,8 +319,8 @@ for method in methods:
                      [time_sqz_D_trunc_strong[method][1][positive_vals-1]],
                      "o", color = colors[method])
 
-plt.xlabel(r"$N\chi t$")
[…]

### 2019-04-28 soc_squeezing/collective_paper/collective_spin.tex
added more commentary about the divergence of correlators added better commentary about functional forms for C(t) and D(t)

@@
 expansion gives essentially exact results right up until a sudden and
-drastic departure that can be diagnosed by inspection.
-% todo: comment on diagnosing breakdown by examining individual correlators, rather than squeezing
-This drastic departure occurs because the TST expansion neglects
-high-weight (i.e.~large-$\abs{\v m}$) operators $\hat\S_{\v m}$ whose
-contributions to a Heisenberg operator $\hat\O\p{t}$ of interest
-eventually become non-negligible.  For sufficiently large times $t$,
-these individual contributions may generally be large compared to the
-value of $\bk{\hat\O\p{t}}$, in which case their truncation will yield
-non-physical results.  This breakdown mechanism is also the reason for
-prolonged agreement between the TST expansion and benchmarking methods
-when squeezing the initial state $\ket{-\Z}$ rather than $\ket{\X}$:
-for initial states $\ket{-\Z}$, all initial-time correlators
-$\bk{\hat\S_{\v m}}_{t=0}$ vanish unless $m_+=m_-=0$ (see Appendix
-\ref{sec:initial_conditions}), so there is a substantially smaller
-number of neglected terms with non-zero contribution to the correlator
-$\bk{\hat\O\p{t}}$.
+drastic departure that can be diagnosed by inspection.  The breakdown
+of the TST expansion in Figure \ref{fig:benchmarking} is marked by an
+unphysical squeezing parameter $\xi^2<0$, which occurs due to
+individual correlators taking unphysical values
+$\lvert\bk{\hat\S_{\v n}}\rvert\gtrsim S^{\abs{\v n}}$.  The sudden
+and drastic departure from virtually exact results occurs because the
+TST expansion neglects high-weight (i.e.~large-$\abs{\v m}$) operators
+$\hat\S_{\v m}$ whose contributions to a Heisenberg operator
+$\hat\O\p{t}$ of interest eventually become non-negligible.  For
+sufficiently large times $t$, these individual contributions may
+generally be large compared to the value of $\bk{\hat\O\p{t}}$, in
+which case their truncation will neglect terms that are large compared
+to the actual value of $\bk{\hat\O\p{t}}$.  This breakdown mechanism
+is also the reason for prolonged agreement between the TST expansion
+and benchmarking methods when squeezing the initial state $\ket{-\Z}$
+rather than $\ket{\X}$: for initial states $\ket{-\Z}$, all
+initial-time correlators $\bk{\hat\S_{\v m}}_{t=0}$ vanish unless
+$m_+=m_-=0$ (see Appendix \ref{sec:initial_conditions}), so there is a
+substantially smaller number of neglected terms with non-zero
+contribution to the correlator $\bk{\hat\O\p{t}}$.
 
@@
 
-% todo: add commentary/discussion about growth of C(t)
-
 Figure \ref{fig:multi_time} shows the behavior of $C\p{t}$ and
@@
 decay, excitation, and dephasing of individual spins at rates
-$\gamma_-=\gamma_+=\gamma_\z=100\chi$.  In the case of TAT and TNT,
-after an initial ramp-up period these correlators grow exponentially
-as $\sim\exp\p{a\chi t}$ with $a\sim N$.  The growth of these
-correlators is definitively sub-exponential for OAT.  In the case of
-unitary time evolution under OAT in particular, to an excellent
-approximation we find that $D\p{t}=D\p{0}+\p{b\chi t}^2$ with
-$b=N+O\p{1}$.  As may be expected, the growth rates of $C\p{t}$ and
-$D\p{t}$ are generally suppressed by decoherence.  Figure
-\ref{fig:multi_time} serves as an example for the type of behavior
-that is accessible at short times with the TST expansion.  These
-examples are straightforward extend to equilibrium settings and
-spin-boson systems.
+$\gamma_-=\gamma_+=\gamma_\z=100\chi$.  In the case of unitary
+evolution under OAT, we find that to an excellent approximation
+$\abs{C\p{t}}$ takes the functional form
+$f\p{t}=f\p{0}+aN\chi t+\p{bN\chi t}^2$ with $a\sim b\sim 1$, and with
+a virtually perfect fit $D\p{t}=D\p{0}+\p{\sp{N+1}\chi t}^2$.  For
+unitary evolution under TAT and TNT, we find that to an excellent
+approximation both $\abs{C\p{t}}$ and $D\p{t}$ take the functional
+form $f\p{t}=f\p{0}+a\sp{\exp\p{bN\chi t}-1}$ with $a\sim b\sim 1$.
+As may be expected, the growth of $C\p{t}$ and $D\p{t}$ is generally
+suppressed by decoherence.  Figure \ref{fig:multi_time} serves as an
+example for the type of behavior that is accessible at short times
+with the TST expansion.  These examples are straightforward extend to
+equilibrium settings and spin-boson systems.
 
@@
   \f{d}{dt}\p{S_\z^2}
-  &= S\p{\gamma_++\gamma_-} + 2\p{\gamma_+-\gamma_-}\p{S-\f12} S_\z
+  &= S\p{\gamma_++\gamma_-} + 2\p{S-\f12}\p{\gamma_+-\gamma_-} S_\z
   - 2 \p{\gamma_++\gamma_-} S_\z^2.
@@
 \begin{align}
-  \bk{S_\z}
-  &= S\p{\f{\gamma_+-\gamma_-}{\gamma_++\gamma_-}}
-  \p{1-e^{-\p{\gamma_+ + \gamma_-} t}}, \\
-  \bk{S_\z^2} &= S \sp{\f12 + \p{S-\f12} \f{\bk{S_\z}^2}{S^2}}.
+  \bk{S_\z} = S\p{\f{\gamma_+-\gamma_-}{\gamma_++\gamma_-}}
+  \p{1-e^{-\p{\gamma_+ + \gamma_-} t}},
+  &&
+  \bk{S_\z^2} = S \sp{\f12 + \p{S-\f12} \f{\bk{S_\z}^2}{S^2}}.
   \label{eq:Sz_OAT}

### 2019-05-01 soc_squeezing/collective_paper/collective_spin.tex
changed a legend and the order in which some lines are drawn

@@
     (dots), and the TST expansion in Eq.~\eqref{eq:TST} with $M=35$
-    (darker dashed lines).  Solid circles mark the times at which the
-    TST expansion gives an unphysical result with $\xi^2<0$.}
+    (dashed lines).  Solid circles mark the times at which the TST
+    expansion gives an unphysical result with $\xi^2<0$.}
   \label{fig:benchmarking}

### 2019-05-01 soc_squeezing/scripts/plot_squeezing.py

@@ -3,7 +3,6 @@
 import os, sys
 import numpy as np
 import matplotlib.pyplot as plt
-import colorsys
 
 from scipy.integrate import solve_ivp
 
@@ -277,6 +276,8 @@ plt.legend(loc = "best", bbox_to_anchor = (0.5, 0.5, 0.5, 0.5))
 plt.tight_layout()
 if save: plt.savefig(fig_dir + "unitary.pdf")
 
+leg_handles, leg_labels = plt.gca().get_legend_handles_labels()
+
 
 ### evolution with weak decoherence
 
@@ -284,7 +285,7 @@ plt.figure(figsize = figsize)
 for method in methods:
     linestyle = "-" if method == OAT else "."
     plt.semilogy(time_sqz_D_exact[method][0],
-                 time_sqz_D_exact[method][1], linestyle, label = method,
+                 time_sqz_D_exact[method][1], linestyle,
                  color = colors[method], markersize = trajectory_marker_size)
     positive_vals = positive(time_sqz_D_trunc[method][1])
     plt.semilogy(time_sqz_D_trunc[method][0][:positive_vals],
@@ -301,7 +302,7 @@ plt.ylabel(r"Squeezing ($\xi^2$)")
 plt.xlim(0, max_plot_time)
 plt.ylim(*ylims(time_sqz_D_exact,time_sqz_D_trunc))
 
-plt.legend(loc = "best", numpoints = 3)
+plt.legend(leg_handles, leg_labels, loc = "best")
 plt.tight_layout()
 if save: plt.savefig(fig_dir + "decoherence_weak.pdf")
 
@@ -313,7 +314,7 @@ for method in methods:
     positive_vals = positive(time_sqz_D_trunc_strong[method][1])
     plt.semilogy(time_sqz_D_trunc_strong[method][0][:positive_vals],
                  time_sqz_D_trunc_strong[method][1][:positive_vals],
-                 color = colors[method], label = method)
+                 color = colors[method])
     if positive_vals < len(times):
         plt.semilogy([time_sqz_D_trunc_strong[method][0][positive_vals-1]],
                      [time_sqz_D_trunc_strong[method][1][positive_vals-1]],
@@ -325,7 +326,7 @@ plt.ylabel(r"Squeezing ($\xi^2$)")
 strong_time_lim_idx = positive(time_sqz_D_trunc_strong[TAT][1])
[…]

### 2019-05-21 soc_squeezing/squeezing_paper/reviews/referee_response.tex
added soc squeezing paper in the middle of review last thing worked on: Appendix D in the main text

\documentclass[reprint,onecolumn,12pt]{revtex4-2}

%%% linking references
\usepackage{hyperref}
\hypersetup{
  colorlinks=true,
  linkcolor=blue,
  urlcolor=cyan,
}

%%% physics and math packages
\usepackage{physics} % general physics package
\usepackage{amssymb} % math fonts and symbols
\usepackage{bm} % for making math symbols bold
\usepackage{braket} % for nice brackets (e.g. |X>)
\usepackage{mathtools} % for \coloneqq
\usepackage{graphicx} % for figures
\graphicspath{{./figures/}} % set path for all figures

%%% shorthands to use inside math environments
\renewcommand{\t}{\text} % text in math mode
\newcommand{\f}[2]{\dfrac{#1}{#2}} % shorthand for fractions
\newcommand{\p}[1]{\left(#1\right)} % parenthesis
\renewcommand{\sp}[1]{\left[#1\right]} % square parenthesis
\renewcommand{\set}[1]{\left\{#1\right\}} % square parenthesis
\renewcommand{\c}{\cdot} % inner product
\renewcommand{\d}{\text{d}} % integration measure, e.g. d^3 x
\renewcommand{\v}{\bm} % bold symbols (e.g. for vectors)
\newcommand{\uv}[1]{\hat{\bm #1}} % unit vectors, or vector operators
\newcommand{\bk}{\Braket}
\renewcommand{\ket}{\Ket}
\renewcommand{\bra}{\Bra}

\newcommand{\B}{\mathcal{B}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\G}{\mathcal{G}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\J}{\mathcal{J}}
\renewcommand{\L}{\mathcal{L}}
\renewcommand{\O}{\mathcal{O}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\Q}{\mathcal{Q}}
\renewcommand{\S}{\mathcal{S}}

\newcommand{\z}{\text{z}}
\newcommand{\x}{\text{x}}
\newcommand{\y}{\text{y}}

\newcommand{\up}{\uparrow}
\newcommand{\dn}{\downarrow}

\usepackage{enumitem}
\setlist[enumerate]{leftmargin=*}
\setlist[itemize]{leftmargin=*}
\usepackage[dvipsnames]{xcolor}
\newcommand{\blue}[1]{{\color{blue} #1}}
\newcommand{\red}[1]{{\color{red} #1}}
\newcommand{\green}[1]{{\color{ForestGreen} #1}}

\begin{document}

\section*{Response to referee}

We thank the referee for taking the time to carefully read our
manuscript, and for providing thoughtful and constructive feedback.
Below, we address the comments and questions brought up by the
referee.  We hope that the referee finds our responses and revisions
satisfactory, deeming our manuscript acceptable for publication in
Physical Review X.

Excerpts from the referee reports are written in \blue{blue}, excerpts
from our original manuscript are written in \red{red}, and excerpts
from the revised version of our manuscript is written in
\green{green}.  All page, equation, and reference numbers generally
refer to those in our original manuscript, unless explicitly stated
otherwise.

\begin{enumerate}
\item Concerning:

  \blue{The main result of the manuscript is the theoretical
    development and characterization of a new protocol for the
    generation of spin squeezing in optical atomic clocks, to be
    employed for boosting the precision of clock experiments beyond
    the standard quantum limit. The topic, encompassing AMO physics
    and quantum physics, is certainly highly interesting. The
    implementation of these ideas in an experiment would constitute a
    very important advancement in the broad field of quantum sensing
    and metrology, as the application of nonclassical states for
    quantum enhancement of the measurement precision is the subject of
    major efforts in quantum technologies nowadays.}

  \blue{The proposal is new, although it is substantially based on
    previous theoretical and experimental works by the groups led by
    the two senior authors (J. Ye and A. M. Rey). On the theoretical
    side, the amount of technological novelty seems limited, as the
    authors are not developing radically new theoretical tools or
    approaches. What is new is the idea that the ``ingredients'' which
    have already been demonstrated can be mixed up, with the proper
    protocol, to generate substantial, useful entanglement. On the
    experimental side, the implementation seems feasible and the
    consequences would be highly relevant, as the promised levels of
    spin squeezing have been never reported yet in any experiment of
    the atomic clock kind. Balancing these considerations, I find the
    subject overall appropriate for PRX, mostly in view of the
    importance of an eventual experimental demonstration of the
    proposed protocol.}

  \blue{The paper is written very well and the level of presentation
    fits the target of PRX, with a first paragraph that serves as a
    quite useful introduction for non-specialists. In the same spirit,
    I have appreciated the decision of the authors to focus on the
    one-axis twisting (OAT) protocol, as it is less effective but more
    pedagogical than the two-axis twisting (TAT) protocol they also
    consider, leaving the latter almost completely in the
    appendixes. Although the interested reader will be forced to read
    the appendixes to understand how the TAT works, I think that this
    choice is positive overall. With the more technical parts moved to
    the appendixes and the general physics presented in the main text,
    the paper seems suitable for a generic audience.}

  \blue{So, on the side of presentation and significance, I find the
    manuscript adequate for PRX.}

  \blue{Before giving a final assessment on the actual validity of the
    manuscript, I invite the authors to provide clarifications on the
    following points:}

  We thank the referee for summarizing the nature and impact of our
  work, deeming the subject of our manuscript ``appropriate for PRX'',
  and noting that our manuscript is ``written very well'' with a
  ``presentation [that] fits the target of PRX''.  We hope that the
  revisions to our manuscript outlined below adequately address any
  questions and reservations that the referee may have, deeming our
  revised manuscript acceptable for publication in PRX.

\item Concerning:

  \blue{1) The authors always consider the case of a homogeneously
    trapped atomic gas, whereas in current 3D optical clock
    implementations there is a significant quasi-harmonic trapping
    induced by the focused magic-optical-lattice beams. Could the
    authors comment on the relevance of this effect for their spin
    squeezing protocol? Since the squeezing arises from the dephasing
    of spatially delocalized wavefunctions, the trap, modifying those
    wavefunctions, might introduce unwanted effects. For instance, the
    spectrum of the B term in the Hamiltonian would change with
    respect to the simple Cos form of Eq. (2): with the upper bound of
    the spectrum upshifted by the harmonic potential energy, the
    (global) interaction gap U would be shrinked, potentially leading
    to a mixing of different Dicke manifolds. In addition, the lattice
    filling f will be continuously changing all over the trap,
    diminishing in the outer regions and potentially causing problems
    for the weak SOC limit assumption. It is not clear to me whether a
    LDA approach would be sufficient to demonstrate the efficiency of
    the protocol in the presence of the trap, as the wavefunctions are
    delocalized. If yes, which assumptions have to be made on the
    system?}

\item Concerning:

  \blue{2) The authors always consider the effects of purely elastic
    interactions. However, in previous works it was shown that
    collisions involving two-electron atoms in the 1S0 and 3P0 states
    can have a substantial inelastic character, with sizable atom loss
    rates. The authors should comment on the influence of such
    processes on their protocol. Can they provide estimates on the
    reduction of spin squeezing when these processes are considered?
    Of course, this effect could be circumvented by considering the
    alternative squeezing protocol relying on nuclear spin states and
    Raman transitions, very briefly mentioned in the last paragraph
    before the conclusions.  Regarding the latter point, it would be
    interesting to know something more about the transfer of
    entanglement from one degree of freedom (nuclear spin) to the
    other (electronic state).}

\item Concerning:

  \blue{3) What happens after the squeezing protocol is never
    mentioned. Even briefly (for the sake of non-experts), the authors
    should explain how the ensuing clock interrogation protocol would
    be implemented. I guess that it would be carried out with a simple
    increase of the lattice depth to stop the effect of the squeezing
    Hamiltonian, followed by a plain clock interrogation sequence in
    the Lamb-Dicke regime. Is this enough or are there additional
    subtleties to be discussed?}

  We thank the referee for their careful consideration of how our
  squeezing protocol would be used in practice.  Indeed, the
  accessibility of our work would benefit greatly from a discussion of
  how to interface our squeezing protocol with an ensuing clock
  interrogation protocol.  To address this point, we have added a
  written a new appendix for our manuscript (Appendix D, on page 10 of
  the revised manuscript):
  % todo: make sure appendix and page numbers are correct

  \green{[Insert Appendix D here, once we finalize it.]}

  We refer readers to the new appendix at the end of the caption to
  Fig.~1 on page 3, adding the text ``\green{(see Appendix D)}''; as
  well as after Eq.~(5) on page 4, replacing the text:

  \red{Note that the above protocol concerns only the preparation of a
    spin-squeezed state, which ideally would be used as an input state
    for a follow-up clock in- terrogation protocol. SOC can be turned
    off during the latter protocol by properly re-aligning the clock
    laser.}

  by:

  \green{The above protocol concerns only the preparation of a
    spin-squeezed state, which would then be used as an input state
    for a follow-up clock interrogation protocol without SOC.  We
    discuss subtleties of interfacing the squeezing protocol with a
    subsequent clock interrogation protocol in Appendix D.}

  We have also added two sentences to the main text in order to make
  readers aware of the pulsed drive TAT protocol.  Specifically, we
  have changed the text (immediately preceding section III on page 4):

  \red{... TAT remains approximately constant until reaching
    Heisenberg-limited amount of spin squeezing with
    $\xi^2_{\t{opt}}\sim N^{-1}$[37].  Following the prescription in
    Ref.~[38], ...}

  which now reads:

  \green{... TAT remains approximately constant until reaching
    Heisenberg-limited amount of spin squeezing with
    $\xi^2_{\t{opt}}\sim N^{-1}$[17, 37].}

  \green{There are two general strategies for converting OAT into TAT:
    by use of a pulsed[38] or continuous[39] drive protocol.  For
    simplicity, we consider the latter in our work.  Following the
    prescription in Ref.~[39], ...}
  % TODO: make sure reference numbers are correct

\item Concerning:

  \blue{4) It is not completely clear whether the results of Fig.~4 in
    the presence of decoherence have been obtained for a fixed
    evolution time (the authors mention a 1s timescale in the main
    text) or they represent the best spin squeezing values reached at
    an optimal squeezing time? Is there still an optimal squeezing
    time in the presence of decoherence or a different, maybe
    asymptotic, evolution?}

  We thank the referee for finding an ambiguity in our presentation of
  the results in Fig.~4.  To clarify that optimal squeezing occurs at
  a particular time that depends on system and decoherence parameters,
  we have modified the caption of Fig.~4 (on page 6), such that the
  text:

  \red{{\bf Squeezing with decoherence} ... The time scales for
    optimal squeezing with decoherence are generally no greater than
    those in Fig.~3.}

  now reads:

  \green{{\bf Optimal squeezing with decoherence} ... Optimal
    squeezing times in the presence of decoherence are generally
    smaller than the corresponding times shown in Fig.~3, as
    decoherence generally degrades squeezing before it reaches the
    decoherence-free maximum.  Sample plots of squeezing over time for
    particular choices of lattice size ($\ell$) and depth
    ($V_0/E_{\t{R}}$) are provided in the Supplementary
    Material[37].}
  % TODO: make sure reference number is correct

  Note that Fig.~3 shows the optimal squeezing time as a function of
  lattice depth ($V_0/E_{\t{R}}$) and system size ($\ell$).  We hope
  that these changes clarify the fact that optimal squeezing generally
  occurs at a particular time that depends on system parameters.

\item Concerning:

  \blue{5) A final minor point. When the authors indicate that the
    motional degrees of freedom are frozen (for instance, in the new
    paragraph after Eq. 5), it would be quite helpful (for readers who
    are not familiar with previous works) to stress again that
    localization is in momentum space, while the wavefunctions in real
    space are extended.}

  We thank the referee for pointing out a potential source of
  confusion from our use of language when we say ``frozen motional
  degrees of freedom''.  Indeed, the atoms in our setup may still
  tunnel from site to site; these atoms are only ``frozen'' in
  momentum space.  In order to clarify this point, we revised all
  references in our manuscript to frozen motional degrees of freedom,
  in particular changing the following text in the caption to Fig.~1
  (on page 3):

  \red{({\bf c}) If interactions are sufficiently weak, all motional
    degrees of freedom are frozen and atoms are effectively pinned to
    fixed quasi-momentum modes $\v q$.}

  to:

  \green{({\bf c}) If interactions are sufficiently weak, all motional
    degrees of freedom become frozen in momentum space, with atoms
    effectively pinned to fixed quasi-momentum modes $\v q$.}

  and adding a parenthetical to the following text after Eq.~(5) on
  page 4:

  \red{... one cannot assume frozen motional degrees of freedom and
    map the Fermi-Hubbard model onto a spin model.}

  which now reads:

  \green{... one cannot assume frozen motional degrees of freedom
    (i.e.~with atoms pinned to fixed quasi-momentum modes) and map the
    Fermi-Hubbard model to a spin model.}

\end{enumerate}

\end{document}

### 2019-05-21 soc_squeezing/squeezing_paper/main.tex
added soc squeezing paper in the middle of review last thing worked on: Appendix D in the main text

\documentclass[aps,prx,superscriptaddress,twocolumn]{revtex4-2}

%%% linking references
\usepackage{hyperref}
\hypersetup{
  colorlinks=true,
  linkcolor=blue,
  urlcolor=cyan,
}

%%% physics and math packages
\usepackage{physics} % general physics package
\usepackage{amssymb} % math fonts and symbols
\usepackage{bm} % for making math symbols bold
\usepackage{braket} % for nice brackets (e.g. |X>)
\usepackage{mathtools} % for \coloneqq
\usepackage{graphicx} % for figures
\graphicspath{{./figures/}} % set path for all figures

%%% shorthands to use inside math environments
\renewcommand{\t}{\text} % text in math mode
\newcommand{\f}[2]{\dfrac{#1}{#2}} % shorthand for fractions
\newcommand{\p}[1]{\left(#1\right)} % parenthesis
\renewcommand{\sp}[1]{\left[#1\right]} % square parenthesis
\renewcommand{\set}[1]{\left\{#1\right\}} % square parenthesis
\renewcommand{\c}{\cdot} % inner product
\renewcommand{\d}{\text{d}} % integration measure, e.g. d^3 x
\renewcommand{\v}{\bm} % bold symbols (e.g. for vectors)
\newcommand{\uv}[1]{\hat{\bm #1}} % unit vectors, or vector operators
\newcommand{\bk}{\Braket}
\renewcommand{\ket}{\Ket}
\renewcommand{\bra}{\Bra}

\newcommand{\B}{\mathcal{B}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\G}{\mathcal{G}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\J}{\mathcal{J}}
\renewcommand{\L}{\mathcal{L}}
\renewcommand{\O}{\mathcal{O}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\Q}{\mathcal{Q}}
\renewcommand{\S}{\mathcal{S}}

\newcommand{\z}{\text{z}}
\newcommand{\x}{\text{x}}
\newcommand{\y}{\text{y}}

\newcommand{\up}{\uparrow}
\newcommand{\dn}{\downarrow}

%%% text for affiliations and contributions
\newcommand{\JILA}{JILA, National Institute of Standards and Technology and Department of Physics, University of Colorado, Boulder, CO, 80309, USA}
\newcommand{\CTQM}{Center for Theory of Quantum Matter, University of Colorado, Boulder, CO, 80309, USA}
\newcommand{\contrib}{\thanks{Authors P.H.~and M.A.P.~contributed equally to this work.}}

\begin{document}

\title{Engineering spin squeezing in a 3D optical lattice with interacting spin-orbit-coupled fermions}

\author{P.~He} \contrib
\author{M.~A.~Perlin} \contrib
\author{S.~R.~Muleady}
\author{R.~J.~Lewis-Swan}
\affiliation{\JILA}
\affiliation{\CTQM}
\author{R.~B.~Hutson}
\author{J.~Ye}
\affiliation{\JILA}
\author{A.~M.~Rey}
\affiliation{\JILA}
\affiliation{\CTQM}

\begin{abstract}
One of the most important tasks in modern quantum science is to coherently control and entangle many-body systems, and to subsequently use these systems to realize powerful quantum technologies such as quantum-enhanced sensors.
However, many-body entangled states are difficult to prepare and preserve since internal dynamics and external noise rapidly degrade any useful entanglement.
Here, we introduce a protocol that counterintuitively exploits inhomogeneities, a typical source of dephasing in a many-body system, in combination with interactions to generate metrologically useful and robust many-body entangled states.
Motivated by current limitations in state-of-the-art three-dimensional (3D) optical lattice clocks (OLCs) operating at quantum degeneracy, we use local interactions in a Hubbard model with spin-orbit coupling to achieve a spin-locking effect.
In addition to prolonging inter-particle spin coherence, spin-locking transforms the dephasing effect of spin-orbit coupling into a collective spin-squeezing process that can be further enhanced by applying a modulated drive.
Our protocol is fully compatible with state-of-the-art 3D OLC interrogation schemes and may be used to improve their sensitivity, which is currently limited by the intrinsic quantum noise of independent atoms.
We demonstrate that even with realistic experimental imperfections, our protocol may generate $\sim10$--$14$ dB of spin squeezing in $\sim1$ second with $\sim10^2$--$10^4$ atoms.
This capability exemplifies a new paradigm of using driven non-equilibrium systems to overcome current limitations in quantum metrology, allowing OLCs to enter a new regime of enhanced sensing with correlated quantum states.
\end{abstract}

% keywords: ultracold atoms, optical lattices, atomic clocks, metrology, spin squeezing

\maketitle

\section{Introduction}

% The understanding and application of few-body quantum mechanics and equilibrium statistical mechanics were integral to the rapid growth in technology throughout the 20th century.
A major frontier of contemporary physics is the understanding of non-equilibrium behaviors of many-body quantum systems, and the application of these behaviors toward the development of novel quantum technologies with untapped capabilities\cite{eisert2015quantum}.
To this end, ultracold atomic, molecular, and optical systems are ideal platforms for studying unexplored regimes of many-body physics due to their clean preparation and readout, high controllability, and long coherence times\cite{bloch2008manybody, gross2017quantum}.
The exquisite capabilities of these systems have pushed the frontiers of metrology, quantum simulation, and quantum information science.

Optical lattice clocks in particular have seen some of the most impressive developments in recent years, reaching record levels of precision ($\sim 3\times 10^{-19}$)\cite{campbell2017fermidegenerate, marti2018imaging} and accuracy ($\sim 1\times 10^{-18}$)\cite{bloom2014optical, mcgrew2018atomic}.
These advancements required important breakthroughs, including the capability to cool and trap fermionic alkaline-earth atoms in spin-insensitive potentials\cite{takamoto2003spectroscopy, barber2006direct, ye2008quantum}; the development of ultra-coherent lasers\cite{kessler2012sub40mhzlinewidth, cole2013tenfold, matei2017mu} to fully exploit an ultra-narrow clock transition\cite{ludlow2015optical}; and, more recently, the preparation of a quantum degenerate gas in a three-dimensional (3D) optical lattice\cite{campbell2017fermidegenerate, marti2018imaging, goban2018emergence}.
Nonetheless, all improvements in sensing capabilities to date have been based on single-particle control of internal atomic degrees of freedom.
Such strategies will eventually have diminishing returns due to practical difficulties in (i) suppressing decoherence from external (motional) degrees of freedom, and (ii) interrogating more particles without additional systematic errors from interactions\cite{martin2013quantum, ludlow2015optical, marti2018imaging}.

Pushing beyond the current independent-particle paradigm requires leveraging many-body quantum correlations.
Entangled states such as spin-squeezed states\cite{kitagawa1993squeezed, wineland1992spin, ma2011quantum} can enhance measurement sensitivity, i.e.~the uncertainty $\Delta\theta$ in the estimation of a parameter $\theta$, below the standard quantum limit $\Delta\theta\sim1/\sqrt{N}$ for $N$ uncorrelated particles\cite{itano1993quantum, degen2017quantum}.
The major challenge for progress in this direction is that generating entanglement requires interactions, which are generally undesirable because they degrade atomic coherence, thereby limiting clock performance\cite{swallows2011suppression, martin2013quantum, rey2014probing, ludlow2011coldcollisionshift, lemke2011wave, ludlow2015optical}.
In fact, the most precise and accurate optical lattice clocks were designed to operate with fermionic atoms in identical nuclear and electronic states in order to suppress collisional decoherence\cite{campbell2009probing, swallows2011suppression, campbell2017fermidegenerate}, as identical fermions cannot interact via the otherwise dominant ($s$-wave) collisions at ultracold temperatures.
However, an initially spin-polarized Fermi gas still exhibits interactions at later times due to spin-orbit coupling (SOC) that is induced by the laser that drives the clock transition (i.e.~the ``clock laser'')\cite{wall2016synthetic, kolkowitz2016spinorbitcoupled, livi2016synthetic, bromley2018dynamics}.
Specifically, the momentum kick imparted by this laser imprints a position-dependent phase that induces inhomogenous spin precession and generates spin dephasing, thereby making atoms distinguishable and vulnerable to collisions.
While a deep lattice can suppress SOC, it also intensifies the light scattering which currently limits the coherence time of the clock\cite{dorscher2018latticeinduced, goban2018emergence, hutson2019engineering}.

In this work, we describe a scheme that can potentially lead to metrological advances in state-of-the-art optical lattice clocks through direct use of quantum entanglement by harnessing the interplay between nominally undesirable collisions and SOC.
This scheme is made possible in the weak SOC regime by the formation of an interaction-energy gap that suppresses the SOC-induced population transfer from the exchange-symmetric Dicke manifold (spanned by spin-polarized and thus non-interacting states) to the remainder of Hilbert space.
In this regime, the non-equilibrium dynamical processes induced by SOC result in all-to-all collective spin interactions that naturally generate spin-squeezed states in a manner robust to realistic experimental imperfections.
To generate spin squeezing, our protocol only requires control over (i) the orientation of the clock laser and (ii) the optical lattice depth.
These controls are straightforward to incorporate into current 3D clock interrogation sequences without sacrificing atom numbers or coherence times.
Additionally, we show that by applying a modulated drive from the clock laser, one can further prepare states that saturate the Heisenberg limit $\Delta\theta \sim 1/N$ for phase sensitivity\cite{kitagawa1993squeezed, ma2011quantum, degen2017quantum}.

\section{From the Fermi-Hubbard model to one-axis and two-axis twisting}
\label{sec:theory}

We consider $N$ fermionic atoms with two spin states (labeled $\up$ and $\dn$) trapped in a 3D optical lattice.
In this discussion, these spin states are associated with the two electronic states of a nuclear-spin-polarized gas.
At sufficiently low temperatures, atoms  occupy the lowest Bloch band of the lattice and interact only through $s$-wave collisions.
A schematic of this system is provided in Fig.~\ref{fig:protocol_schematic}(a), where tight confinement prevents motion along the vertical direction ($z$), effectively forming a stack of independent 2D lattices.
For simplicity and without loss of generality, however, we first consider the case when tunneling can only occur along one direction, $x$, and thus model the system as living in  one dimension.

\begin{figure*}
\centering
\includegraphics[width=0.6\textwidth]{protocol_schematic.pdf}
\caption{{\bf Schematic of the setup for spin squeezing}.
({\bf a}) We consider $N$ fermionic atoms with two (pseudo-)spin components, represented by red and blue spheres, trapped in the ground band of an optical lattice (shown in 2D for the sake of presentation).
Atoms tunnel to neighboring sites at a rate $J$ and experience on-site interactions with strength $U$.
An external laser carrying a position dependent phase $e^{i\v k_{\t{L}}\cdot\v r}$ couples the spin states of the atoms.
({\bf b}) After a gauge transformation, different spin states exhibit different dispersion relations with a relative phase $\phi=k_{\text{L}} a$, where $a$ is the lattice spacing.
The external laser couples spin states with identical quasi-momenta $q$ in the gauge-transformed frame.
({\bf c}) If interactions are sufficiently weak, all motional degrees of freedom become frozen in momentum space, with atoms effectively pinned to fixed quasi-momentum modes $\v q$.
The dynamics on the frozen $\v q$-space lattice can then be mapped to a spin model in which collisional interactions correspond to a uniform, all-to-all ferromagnetic Heisenberg Hamiltonian with strength $U/L$, where $L$ is the total number of lattice sites.
({\bf d}) The spin dependence of the dispersion relation is captured by a mode-dependent axial field $B_q$ that generates inhomogeneous spin precession.
This axial field couples exchange-symmetric many-body Dicke states with total spin $S=N/2$ to spin-wave states with $S=N/2-1$.
The all-to-all interaction opens an energy gap $fU$ (with $f=N/L$ the filling fraction of spatial modes) between the Dicke states and the spin-wave states, which forbids population transfer between them in the weak-field limit.
({\bf e}) To generate spin squeezing via one-axis twisting, we initialize a product state with all spins polarized in $-z$ (i.e.~in $\ket{\dn}$), and apply a fast external laser pulse to rotate all spins into $x$.
We then let atoms freely evolve for a variable time $t$ (with a spin-echo pulse), after which the amount of spin squeezing can be determined experimentally from global spin measurements.
The spin-squeezed state can be used for a follow-up clock interrogation protocol (see Appendix \ref{sec:clock_interrogation}).
}
\label{fig:protocol_schematic}
\end{figure*}

An external laser with Rabi frequency $\Omega$ and wavenumber $k_{\t{L}}$ along the tunneling axis resonantly couples atoms' internal states through the Hamiltonian $\hat H_{\t{laser}}/\hbar = \sum_j \Omega e^{ik_{\t{L}}x_j} \hat c_{j,\up}^\dag \hat c_{j,\dn} + \t{h.c.}$, where $\hat c_{j\alpha}$ is a fermionic annihilation operator for an atom on site $j$ with internal state $\alpha\in\set{\up,\dn}$ and $x_j$ is the position of site $j$.
This laser imprints a position-dependent phase that equates to a momentum kick $k_{\t{L}}$ when an atom changes internal states by absorbing or emitting a photon, thereby generating spin-orbit coupling\cite{wall2016synthetic, livi2016synthetic}.
After absorbing the position dependence of the laser Hamiltonian into fermionic operators through the gauge transformation $\hat c_{j,\up}\to e^{ik_{\t{L}} x_j}\hat c_{j,\up}$, which makes $\hat H_{\t{laser}}$ spatially homogeneous, the atoms are well-described in the tight-binding limit by the Fermi-Hubbard Hamiltonian\cite{esslinger2010fermihubbard}
\begin{align}
  \hat H_{\text{FH}}/\hbar
  &= -J \sum_j \p{e^{i\phi} \hat c_{j,\up}^\dag \hat c_{j+1,\up}
  + \hat c_{j,\dn}^\dag \hat c_{j+1,\dn} + \text{h.c.}} \notag \\
  &\qquad + U \sum_j \hat n_{j,\up} \hat n_{j,\dn},
\label{eq:FermiHubbard}
\end{align}
where $J$ is the nearest-neighbor tunneling rate; the SOC angle $\phi\equiv k_{\t{L}}a$ determines the phase gained by spin-up atoms upon tunneling from site $j$ to site $j+1$ (in the gauge-transformed frame) with lattice spacing $a=x_{j+1}-x_j$; $U$ is the on-site interaction energy of two atoms; and $\hat n_{j\alpha}\equiv\hat c_{j\alpha}^\dag \hat c_{j\alpha}$ is a number operator.

The Fermi-Hubbard Hamiltonian can be re-written in the quasi-momentum basis with annihiliation operators $\hat c_{q\alpha}\equiv L^{-1/2}\sum_j e^{-iqx_j}\hat c_{j\alpha}$, where $q$ is a quasi-momentum and $L$ is the total number of lattice sites.
In this basis, the single-particle Hamiltonian exhibits shifted dispersion relations that signify spin-orbit coupling [see Fig.~\ref{fig:protocol_schematic}(b)]:
\begin{align}
  \hat H_{\t{FH,single}}/\hbar
  = -2J\sum_q\sp{\cos(qa + \phi) \hat n_{q,\up}
  + \cos(qa) \hat n_{q,\dn}}.
\label{eq:FermiHubbards}
\end{align}
When $U\lesssim J$, interaction energies are too weak for collisions to change the occupancies of single-particle quasi-momentum modes.
Atoms are then pinned to these modes, which form a lattice in quasi-momentum space [see Fig.~\ref{fig:protocol_schematic}(c)]\cite{bromley2018dynamics}.
In this strong-tunneling limit, the Fermi-Hubbard Hamiltonian [Eqn.~\eqref{eq:FermiHubbard}] can be mapped to a spin-$1/2$ system with a collective ferromagnetic Heisenberg interaction and an inhomogeneous axial field, given by\cite{martin2013quantum,rey2014probing, bromley2018dynamics}
\begin{align}
  \hat H_{\t{spin}}/\hbar
  = -\frac{U}{L} \hat{\v S} \c \hat{\v S}
  - \sum_q B_q \hat s_q^\z,
\label{eq:SpinModel}
\end{align}
where $\hat{\v S}=\sum_q\hat{\v s}_q$ is a collective spin operator; $\hat{\v s}_q$ is a spin-1/2 operator for mode $q$ with components $\hat{s}_q^{j=\x,\y,\z}\equiv\frac12\sum_{\alpha,\beta}\hat c_{q\alpha}^\dag\sigma^j_{\alpha\beta}\hat c_{q\beta}$ defined in terms of the Pauli matrices $\sigma^{j=\x,\y,\z}$; the sums over $q$ run over all occupied quasi-momentum modes; and $B_q \equiv -4J\sin(qa+\phi/2)\sin(\phi/2)$ is the SOC-induced axial field.

On its own, the collective Heisenberg term ($\sim\hat{\v S}\c\hat{\v S}$) in Eqn.~\eqref{eq:SpinModel} opens an energy gap $fU$, with $f\equiv N/L$ the filling fraction of spatial modes, between the collective Dicke states $\ket{S=N/2, M_S}$ and the remainder of Hilbert space\cite{rey2008manybody, martin2013quantum, norcia2018cavitymediated} with $S<N/2$.
Here $S$ and $M_S$ respectively label the eigenvalues of the collective spin operators $\hat{\v S}\c\hat{\v S}$ and $\hat{S}_\z$, with eigenvalues $S(S+1)$ for non-negative $S\in\set{N/2,N/2-1,\cdots}$ and $M_S\in\set{-S,-S+1,\cdots,S}$.
The axial field $B_q$ generally couples states within the Dicke manifold to states outside it.
In the weak SOC limit (i.e.~$B_q\ll fU$), however, the interaction energy gap suppresses population transfer between states with different total spin $S$ [see Fig.~\ref{fig:protocol_schematic}(d)].
%In this weak SOC limit (i.e.~of all $B_q/fU\ll1$), we can treat the axial field perturbatively and perform a Schrieffer-Wolff transformation to determine its effective action on states restricted to the Dicke manifold (see Appendix \ref{sec:derivation_OAT}).
In this regime, the virtual occupation of states outside the Dicke manifold can be accounted for perturbatively.
The symmetries of SOC as expressed in Eqn.~\eqref{eq:SpinModel} dictate that this treatment should yield powers of $\hat S_\z$ when projected onto the collective Dicke manifold at higher orders in perturbation theory.
At second order in perturbation theory (see Appendix \ref{sec:derivation_OAT}), we thus find that SOC effectively yields a one-axis twisting (OAT) model widely known to generate squeezing dynamics\cite{kitagawa1993squeezed, ma2011quantum}:
\begin{align}
  \hat H_{\t{eff}}/\hbar
  = -\frac{U}{L} \hat{\v S} \c \hat{\v S}
  - \overline B \hat S_\z + \chi \hat S_\z^2,
  &&
  \chi \equiv \f{\widetilde B^2}{(N-1)fU},
  \label{eq:H_eff}
\end{align}
where $\overline B\equiv\sum_q B_q/N$ is the mean and $\widetilde B^2\equiv\sum_q\p{B_q-\overline B}^2/N$ the variance of the axial field.
The effect of the $\sim\v S\c\v S$ term is to generate a relative phase between states with different total spin $S$ and thus has no effect on dynamics  restricted to a fixed $S$.
Note also that  the collective spin rotation from $\overline B \hat S_\z$  can be eliminated by going into a rotating frame or by using a spin echo.

The entire protocol for preparing a squeezed state via OAT, sketched out in Fig.~\ref{fig:protocol_schematic}(e), reduces to a standard Ramsey protocol with a spin echo: after initially preparing a spin-down (i.e.~$-\uv z$) polarized sample of ultracold atoms populating the lowest Bloch band of a lattice, a fast $\pi/2$ pulse is applied with the clock laser to rotate all spin vectors into $+\uv x$.
The atoms then freely evolve for a variable time $t$ (possibly with spin-echo $\pi$-pulses), after which the amount of metrologically useful spin squeezing, measured by the Ramsey squeezing parameter
\begin{align}
  \xi^2 \equiv
  \min_\theta \braket{\t{var}(\hat S^\perp_\theta)}
  \times N/\abs*{\braket{\hat{\v S}}}^2,
\end{align}
can be determined experimentally from global spin measurements.
Here $\braket{\hat{\v S}}$ is the mean collective-spin vector and $\braket{\t{var}(\hat S^\perp_\theta)}$ is the variance of spin measurements along an axis orthogonal to $\braket{\hat{\v S}}$, parameterized by the angle $\theta\in[0,2\pi)$.
The above protocol concerns only the preparation of a spin-squeezed state, which would then be used as an input state for a follow-up clock interrogation protocol without SOC.
We discuss subtleties of interfacing the squeezing protocol with a subsequent clock interrogation protocol in Appendix \ref{sec:clock_interrogation}.

The validity of the OAT model in Eqn.~\eqref{eq:SpinModel} relies on two key conditions concerning experimental parameter regimes.
First, the on-site interaction energy $U$ should not be much larger in magnitude than the tunneling rate $J$ (clarified below); otherwise, one cannot assume frozen motional degrees of freedom (i.e.~with atoms pinned to fixed quasi-momentum modes) and map the Fermi-Hubbard model to a spin model.
Second, the SOC-induced fields $B_q\sim J\sin(\phi/2)$ should be considerably smaller in magnitude than the interaction energy gap $fU$, as otherwise one cannot perturbatively transform SOC into OAT.
These two conditions can be satisfied by appropriate choices of $U/J$ and the SOC angle $\phi$, which are respectively controlled by tuning the lattice depth and changing the angle between the clock laser and shallow lattice axes [see Fig.~\ref{fig:protocol_schematic}(a)].

\begin{figure*}
\centering
\includegraphics[width=0.8\textwidth]{model_benchmarking.pdf}
\caption{{\bf Benchmarking the spin and one-axis twisting models}.
Comparisons of maximum squeezing (top panels, {\bf a.i} and {\bf b.i}) and optimal squeezing time (lower panels, {\bf a.ii} and {\bf b.ii}) between the Fermi-Hubbard (FH), spin, and one-axis twisting (OAT) models; obtained numerically via the protocol depicted in Fig.~\ref{fig:protocol_schematic}(e) in a 1D lattice with $L=12$ sites.
Results are shown for half filling with $N=12,f\equiv N/L=1$ (left panels, {\bf a.i} and {\bf a.ii}) and filling $f=5/6$ (right panels, {\bf b.i} and {\bf b.ii}) as a function of $U/J$ and the SOC angle $\phi$.
In both cases, the system is initialized in the corresponding ground state.
Insets for both $f=1$ and $f=5/6$ show (in green) regions of the $U$-$\phi$ plane in which both the optimal squeezing (in dB) and the corresponding squeezing time of all three models agree to within 20\%.
At half filling ({\bf a.i} and {\bf a.ii}), mode-changing collisions are suppressed by Pauli blocking, resulting in almost exact agreement between the FH and spin models; both of these models converge onto the OAT model in the gap-protected, weak SOC regime of large $U/J$ and small $\phi$.
The spin and OAT models show similar behavior away from half filling ({\bf b.i} and {\bf b.ii}), but the presence of mode-changing collisions results in their disagreement with the FH model as interactions begin to dominate at larger $U/J$.
Even below half filling, however, the FH exhibits comparable amounts of squeezing to the spin model across a broad range of $U/J$ and $\phi$, albeit at earlier times when $U/J\gtrsim2$.
}
\label{fig:model_benchmarking}
\end{figure*}

We demonstrate the importance of these conditions in Fig.~\ref{fig:model_benchmarking}, where we show numerical results from exact simulations of a 1D system with $L=12$ sites.
Therein, optimal squeezing achievable under unitary dynamics is provided in dB, i.e.~$-10\log_{10}(\xi_{\t{opt}}^2)$, while the time at which this squeezing occurs is provided in units of the nearest-neighbor tunneling time $2\pi/J$.
At $f=1$ atom per lattice site, i.e.~half filling of all atomic states in the lowest Bloch band, the spin model [Eqn.~\eqref{eq:SpinModel}] agrees almost exactly with the Fermi-Hubbard (FH) model [Eqn.~\eqref{eq:FermiHubbard}] up through $U/J=8$. The agreement at half filling ($f=1$) is assisted by Pauli blocking of mode-changing collisions.
Below half filling ($f=5/6$), these two (FH and spin) models show good agreement at $U/J\lesssim2$, while at $U/J\gtrsim2$ mode-changing collisions start to become relevant and invalidate the frozen-mode assumption of the spin model.
Note that we chose filling $f=5/6$ to demonstrate that our protocol should work even in this highly hole-doped case; in practice, optimized experiments are capable of achieving fillings closer to $f=1$\cite{brown2017spinimbalance}.
Interestingly, even with mode-changing collisions the Fermi-Hubbard model exhibits comparable amounts of squeezing to the spin model, and achieves this squeezing in less time.
The spin and OAT models agree in the regime of weak SOC with $\widetilde{B}\sim J\sin(\phi/2)\ll fU$, and exhibit different squeezing behaviors outside this regime as single-particle spin dephasing can no longer be treated as a weak perturbation to the spin-locking  interactions.

The above scheme for OAT achieves optimal spin squeezing that scales as $\xi^2_{\t{opt}}\sim N^{-2/3}$ with minimal intervention, i.e.~a standard Ramsey protocol.
Further improvements upon this scheme can be made by introducing a time-dependent driving field that transforms the OAT Hamiltonian into a two-axis twisting (TAT) one.
While the OAT model initially generates squeezing faster than the TAT model, the squeezing generation rate of OAT (measured in dB per second) falls off with time, while the squeezing generation rate for TAT remains approximately constant until reaching Heisenberg-limited amount of spin squeezing with $\xi^2_{\t{opt}}\sim N^{-1}$\cite{kitagawa1993squeezed, supplementary_material}.

There are two general strategies for converting OAT into TAT: by use of a pulsed\cite{liu2011spin} or continuous\cite{huang2015twoaxis} drive protocol.
For simplicity, we consider the latter in our work.
Following the prescription in Ref.~[\citenum{huang2015twoaxis}], we use the clock laser to apply an amplitude-modulated drive $\hat H_{\t{drive}}(t)/\hbar=\Omega_0\cos(\omega t)\hat{S}_\x$.
If the modulation frequency $\omega$ satisfies $\omega\gg N\chi$ and $\J_0\p{2\Omega_0/\omega}=\pm1/3$, where $\chi$ is the OAT squeezing strength in Eqn.~\eqref{eq:H_eff} and $\J_0$ is the zero-order Bessel function of the first kind, then up to (i) an $\sim\hat{\v S}\c\hat{\v S}$ term that contributes only overall phase factors, and (ii) an $\sim\hat S_\z$ term that can be eliminated with a simple dynamical decoupling pulse sequence (see Appendix \ref{sec:dynamical_decoupling}), the effective Hamiltonian becomes $\hat H_{\t{TAT}}^{(+)}/\hbar = (\chi/3)(\hat S_\z^2-\hat S_\x^2)$ or $\hat H_{\t{TAT}}^{(-)}/\hbar = (\chi/3)(\hat S_\y^2-\hat S_\x^2)$ (see Appendix \ref{sec:derivation_TAT}), which squeezes an initial state polarized along the $y$ or $z$ axis, respectively.

\section{Experimental preparation of a squeezed state and practical considerations}

Here we discuss specific implementations of the above protocols in the state-of-the-art 3D $^{87}$Sr optical lattice clock (OLC).
This system has demonstrated the capability to load a quantum degenerate gas into a 3D lattice at the ``magic wavelength'' ($\lambda_{\t{lattice}}=2a\approx813$ nm) for which both the ground ($^1S_0,\dn$) and first excited ($^3P_0,\up$) electronic states of the atoms experience the same optical potential\cite{campbell2017fermidegenerate}.
Furthermore, the 3D $^{87}$Sr OLC currently operates at sufficiently low temperatures to ensure vanishing population above the lowest Bloch band, such that its dynamics are governed by the Fermi-Hubbard Hamiltonian [Eqn.~\eqref{eq:FermiHubbard}]\cite{esslinger2010fermihubbard}.

An external clock laser with wavelength $\lambda_{\t{L}}\approx698$ nm resonantly interrogates the $^1S_0$ and $^3P_0$ states of the atoms and generates spin-orbit coupling (SOC)\cite{wall2016synthetic}.
While the relative wavelengths of the lattice and clock lasers do not allow for weak SOC along all three lattice axes, weak SOC along two axes can be implemented by, for example, (i) fixing a large lattice depth along the $z$ axis, effectively freezing atomic motion along  $z$, and then (ii) making the clock laser nearly collinear with the $z$ axis, with only a small projection of its wavenumber $\v k_{\t{L}}$ onto the $x$-$y$ plane [see Figure \ref{fig:protocol_schematic}(a)].
The entire 3D OLC then factorizes into an array of independent 2D systems with $N=f\ell^2$ atoms each, where $\ell$ is the number of lattice sites along each axis of the lattice.
As in the 1D case, atoms within the 2D system experience all-to-all interactions, as well as spin-orbit coupling along two directions characterized by SOC angles $\phi_{\x,\y}=k_{\t{L}}^{\x,\y}a$.
Generally speaking, higher-dimensional systems (e.g.~2D vs.~1D) are more desirable because they allow packing more interacting atoms into a fixed system volume, thereby increasing the maximally attainable amount of spin squeezing.

\begin{figure*}
\centering
\includegraphics{optima_coherent.pdf}
\caption{{\bf Optimal squeezing} with one- and two-axis twisting in a 2D section of the 3D $^{87}$Sr optical lattice clock.
({\bf a}) The maximum amount of squeezing depends only on the atom number $N=\ell^2$, where $\ell$ is the number of lattice sites along each axis of the lattice.
While the time scales for squeezing generally depend on several experimental parameters, the time at which maximal squeezing occurs can be minimized at any given lattice depth $V_0$ by choosing SOC angles $\phi$ that saturate $\widetilde{B}/U\approx0.05$, where $\widetilde{B}$ is the variance of the SOC-induced axial field and $U$ is the two-atom on-site interaction energy.
Panels ({\bf b}, {\bf c}) show these minimal squeezing times as a function of the depth $V_0$ and linear size $\ell$ of the lattice.
Lattice depths $V_0$ are normalized to the atomic lattice recoil energy $E_{\t{R}}$, and the upper axis on panels ({\bf b}, {\bf c}) marks values of  $U/J$ at fixed lattice depths.
In general, TAT achieves more squeezing than OAT for any system size, and achieves this squeezing faster for $N\gtrsim500$ atoms.
}
\label{fig:optima_coherent}
\end{figure*}

Figure \ref{fig:optima_coherent} shows, for both OAT and TAT protocols, the maximally attainable amount of spin squeezing and the shortest time at which it occurs as a function of the lattice depth $V_0$ and linear lattice size $\ell$ in a single half-filled 2D layer (i.e.~$f=1,N=\ell^2$) of the 3D OLC.
Atoms are confined along the direction transverse to the 2D layer by a lattice of depth 60 $E_{\t{R}}$, where $E_{\t{R}}$ is the atomic lattice recoil energy.
The maximally attainable amount of spin squeezing by each protocol in Fig.~\ref{fig:optima_coherent} depends only on the atom number $N$, while the shortest attainable time is determined by choosing the largest SOC angles $\phi_\x=\phi_\y\equiv\phi$ which saturate $\widetilde{B}/U\approx0.05$.
We impose this constraint on $\widetilde{B}/U$ to ensure validity of the OAT Hamiltonian perturbatively derived in Appendix \ref{sec:derivation_OAT}\cite{supplementary_material}.

\begin{figure*}
\centering
\includegraphics{optima_decoherence.pdf}
\caption{{\bf Optimal squeezing with decoherence} via one- and two-axis twisting in a 2D section of the 3D $^{87}$Sr optical lattice clock (OLC).
In practice, decoherence due to light scattering  limits the amount of squeezing that is attainable in the the 3D $^{87}$Sr OLC.
Due to growing squeezing times with increasing system size, the maximal squeezing obtainable via OAT saturates past $\ell\approx30$ sites along each axis of the lattice, with $N\approx10^3$ atoms total.
The more favorable size-dependence of TAT time scales,  however, allow for continued squeezing gains through $\ell=100$ ($N=10^4$).
While the OAT results in ({\bf a}) are exact, the TAT results in ({\bf b}) reflect only a lower bound on the maximum squeezing obtainable, albeit one that is likely close (within a few dB) to the actual value.
Optimal squeezing times in the presence of decoherence are generally smaller than the corresponding times shown in Fig.~\ref{fig:optima_coherent}, as decoherence generally degrades squeezing before it reaches the decoherence-free maximum.
Sample plots of squeezing over time for particular choices of lattice size ($\ell$) and depth ($V_0/E_{\t{R}}$) are provided in the Supplementary Material\cite{supplementary_material}.
}
\label{fig:optima_decoherence}
\end{figure*}

Currently, light scattering from the lattice beams induces decoherence of the clock on a time scale of $\sim10$ seconds\cite{goban2018emergence, hutson2019engineering}, which is much shorter than the natural $^3P_0$ lifetime of $\sim160$ seconds (see Appendix \ref{sec:decoherence}).
This limitation imposes significant constraints on achievable spin squeezing, as shown in Figure \ref{fig:optima_decoherence} where the maximal squeezing with spin decay in the OAT case was determined using exact expressions for spin correlators derived in Ref.~[\citenum{foss-feig2013nonequilibrium}], while in the TAT case these correlators were determined by solving Heisenberg equations of motion for collective spin operators (see Appendix \ref{sec:collective_simulation}).
Due to the fast growth of Heisenberg operators in systems with all-to-all interactions, the latter method is not always capable of simulating up to the optimal squeezing time, and thus only provides a lower bound on the maximal squeezing theoretically obtainable via TAT.
The results in Fig.~\ref{fig:optima_decoherence} show that squeezing via OAT saturates with system size around $N\approx10^3$ ($\ell\approx30$), while TAT allows for continued squeezing gains through $N=10^4$ ($\ell=100$).
Even with decoherence, our protocol may realistically generate $\sim10$--$14$ dB of spin squeezing in $\sim1$ second with $\sim10^2$--$10^4$ atoms in a 2D section of the lattice, which is compatible with the atom numbers and interrogation times of state-of-the-art optical lattice clocks\cite{campbell2017fermidegenerate, marti2018imaging}.

The source of decoherence considered above is not fundamental, and can be avoided by, for example, using two nuclear spin levels as spin-1/2 degrees of freedom that are interrogated by far-detuned Raman transitions instead of a direct optical transition\cite{mancini2015observation}.
The strength of SOC for Raman transitions is tunable and, moreover, the lifetimes of ground nuclear spin levels are longer than 100 seconds in the lattice\cite{goban2018emergence}.
In this case, our protocol for preparing a squeezed state would additionally end with a coherent state transfer from nuclear to electronic degrees of freedom in order to retain metrological utility for the atomic clock.

\section{Conclusions}

We have proposed a new protocol to generate spin squeezing in a fermionic 3D optical lattice clock by combining nominally undesirable atomic collisions with spin-orbit coupling.
To our knowledge, this is the first proposal to use atomic collisions for entanglement generation in state-of-the art atomic clocks, opening a path for the practical improvement of world-leading quantum sensors using correlated many-body fermionic states.
Such capability could allow for major improvements in clock sensitivity and bandwidth, enhancing not only traditional timekeeping applications such as measurement standards, navigation (GPS), and telecommunications, but also geodesy and gravitational wave detection, precision tests of fundamental physics, and the search for new physics beyond the standard model\cite{safronova2018search}.

\section{Acknowledgments}

We acknowledge helpful discussions with M.~Norcia, C.~Sanner, and M.~Mamaev.
This work is supported by the Air Force Office of Scientific Research (AFOSR) grant FA9550-18-1-0319; the AFOSR Multidisciplinary University Research Initiative (MURI) grant; the Defense Advanced Research Projects Agency (DARPA) and Army Research Office (ARO) grant W911NF-16-1-0576; the National Science Foundation (NSF) grant PHY-1820885; JILA-NSF grant PFC-173400; and the
National Institute of Standards and Technology (NIST).

\appendix

% make figure numbers include the appendix letter
\renewcommand\thefigure{\thesection\arabic{figure}}

\section{Derivation of the effective one-axis-twisting model}
\label{sec:derivation_OAT}

Suppose we have a Hamiltonian of the form ($\hbar=1$)
\begin{align}
  H = H_0 + V,
\end{align}
with
\begin{align}
  H_0 = - \f{U}{L} \v S\c\v S,
  &&
  V = - \sum_n B_n s_\z^{(n)} + \Omega S_\x,
\end{align}
and we consider $N$-particle states initially in the ground-state manifold $\G_0$ of $H_0$, which have total spin $S=N/2$.
If the largest eigenvalue of $V$ is smaller in magnitude than half of the collective spin gap $NU/L=fU$, i.e.~the energy gap under $H_0$ between $\G_0$ and its orthogonal complement $\E_0$, then we can formally develop a perturbative treatment for the action of $V$ on $\G_0$.
Such a treatment yields an effective Hamiltonian on $\G_0$ of the form $H_{\t{eff}}=\sum_pH_{\t{eff}}^{(p)}$, where $H_{\t{eff}}^{(p)}$ is order $p$ in $V$.
Letting $\P_0$ ($\Q_0$) be a projector onto $\G_0$ ($\E_0$), we define the super-operators $\O$ and $\L$ by
\begin{align}
  \O V &\equiv \P_0 V \Q_0 + \Q_0 V \P_0, \\
  \L V &\equiv \sum_{\alpha,\beta}
  \f{\op{\alpha}\O V \op{\beta}}{E_\alpha-E_\beta},
\end{align}
where $H_0 = \sum_\alpha E_\alpha \op\alpha$.
The first few terms in the expansion of the effective Hamiltonian $H_{\t{eff}}$ are then, as derived in Ref.~[\citenum{bravyi2011schrieffer}],
\begin{align}
  H_{\t{eff}}^{(0)} = \P_0 H_0 \P_0,
  &&
  H_{\t{eff}}^{(1)} = \P_0 V \P_0,
\end{align}
and
\begin{align}
  H_{\t{eff}}^{(2)} = -\f12 \P_0 \sp{\O V,\L V}_- \P_0,
\end{align}
with $\sp{X,Y}_-\equiv XY-YX$.
The zero-order effective Hamiltonian $H_{\t{eff}}^{(0)}=H_0$ within the ground-state manifold.
To calculate $H_{\t{eff}}^{(1)}$, we note that the ground-state manifold $\G_0$ is spanned by the Dicke states
\begin{align}
  \ket{m} \propto S_+^{N/2+m} \ket{\downarrow}^{\otimes n},
  &&
  S_+ \equiv \sum_n s_+^{(n)},
\end{align}
in terms of which we can expand the collective spin-z operator as $S_\z=\sum_m m\op{m}$.
We can likewise expand the collective spin-x operator $S_\x$ in terms of $x$-oriented Dicke states $\ket{m_\x}$ as $S_\x=\sum_m m\op{m_\x}$.
The ground-state projector $\P_0$ onto $\G_0$ can be expanded in either basis as $\P_0=\sum_m\op{m}=\sum_m\op{m_\x}$.
Defining the mean and residual fields
\begin{align}
  \overline B \equiv \f1N \sum_n B_n,
  &&
  b_n \equiv B_n - \overline B,
\end{align}
we can then write
\begin{align}
  V &= -\sum_n \p{b_n+\overline B} s_\z^{(n)} + \Omega S_\x \notag \\
  &= -\sum_n b_n s_\z^{(n)} - \overline B S_\z + \Omega S_\x,
\end{align}
and in turn
\begin{align}
  H_{\t{eff}}^{(1)}
  &= \P_0\p{-\sum_n b_n s_\z^{(n)} - \overline B S_\z + \Omega S_\x} \P_0 \notag \\
  &= -\sum_n b_n \P_0 s_\z^{(n)} \P_0 - \overline B S_\z + \Omega S_\x,
\end{align}
where we used the fact that $\P_0 S_{j=\z,\x} \P_0 = S_j$ within the ground-state manifold.
By construction, the residual fields are mean-zero, i.e.~$\sum_n b_n=0$.
Using the particle-exchange symmetry of the Dicke states, we can therefore expand
\begin{multline}
  \sum_n b_n \P_0 s_\z^{(n)} \P_0
  = \sum_{n,m,m'} b_n \op{m} s_\z^{(n)} \op{m'} \\
  = \sum_n b_n \sum_{m,m'} \op{m} s_\z^{(1)} \op{m'}
  = 0,
\end{multline}
which implies
\begin{align}
  H_{\t{eff}}^{(1)} = - \overline B S_\z + \Omega S_\x.
\end{align}
To calculate the second-order effective Hamiltonian $H_{\t{eff}}^{(2)}$, we let $\B_0\p{\E_0}$ denote an eigenbasis of $H_0$ for the excited subspace $\E_0$, and set the ground-state energy to 0.
We then define the operator
\begin{align}
  \I \equiv \sum_{\ket\alpha\in\B_0\p{\E_0}} \f{\op\alpha}{E_\alpha},
\end{align}
which sums over projections onto excited states with corresponding energetic suppression factors, in terms of which we can write
\begin{align}
  H_{\t{eff}}^{(2)} = -\P_0 V \I V \P_0,
\end{align}
which is simply an operator-level version of the textbook expression for second-order perturbation theory.
The only part of $V$ which is off-diagonal with respect to the ground- and excited-state manifolds $\G_0$ and $\E_0$ is $-\sum_n b_n s_\z^{(n)}$, and the individual spin operators in this remainder can only change the total spin $S$ by at most 1.
It is therefore sufficient to expand $\I$ in a basis for states which span the image of $\G_0$ under all $s_\z^{(n)}$ within the $S=N/2-1$ manifold.
Such a basis is provided by the spin-wave states
\begin{align}
  \ket{mk}
  \propto
  \sum_{n=1}^N e^{2\pi ikn/N} s_+^{(n)} \ket{m-1},
\end{align}
for $k=1,2,\cdots,N-1$\cite{swallows2011suppression}.
Using the fact that all spin-$z$ operators preserve the projection of total spin onto the $z$ axis, we then have that
\begin{multline}
  H_{\t{eff}}^{(2)}
  = -\f1{fU} \sum_{m,k,n,n'} b_n b_{n'}
  \bk{m | s_\z^{(n)} |mk} \\
  \times \bk{mk | s_\z^{(n')} | m},
  \label{eq:general_H_eff_2}
\end{multline}
where the relevant matrix elements between the Dicke states and the spin-wave states are\cite{swallows2011suppression}
\begin{align}
  \bk{m|s_\z^{(n)}|mk}
  = e^{2\pi i k n/N} \sqrt{\f{(N/2)^2-m^2}{N^2 (N-1)}},
\end{align}
which implies
\begin{multline}
  H_{\t{eff}}^{(2)}
  = -\f1{fU} \sum_m \f{(N/2)^2-m^2}{N^2 (N-1)} \op{m} \\
  \times \sum_{k,n,n'} b_n b_{n'} e^{2\pi ik\p{n-n'}/N}.
\end{multline}
Using the fact that $\sum_nb_n=0$, we can expand
\begin{align}
  \sum_{k,n,n'} b_n b_{n'} e^{2\pi ik\p{n-n'}/N}
  &= \sum_{n,n'} b_n b_{n'} \sum_{k=1}^{N-1} e^{2\pi ik\p{n-n'}/N} \notag \\
  &= \sum_{n,n'} b_n b_{n'} \sum_{k=0}^{N-1} e^{2\pi ik\p{n-n'}/N},
\end{align}
where the sum over $k$ vanishes for $n\ne n'$ and equals $N$ when $n=n'$, so
\begin{align}
  \sum_{k,n,n'} b_n b_{n'} e^{2\pi ik\p{n-n'}/N}
  = N^2 \widetilde B^2,
\end{align}
where
\begin{align}
  \widetilde B^2
  \equiv \f1N \sum_n b_n^2 = \f1N \sum_n \p{B_n - \overline B}^2.
  \label{eq:sum_knn}
\end{align}
We therefore have that
\begin{align}
  H_{\t{eff}}^{(2)}
  = -\sum_m \f{\p{N/2}^2-m^2}{\p{N-1}fU}~ \widetilde B^2 \op{m},
\end{align}
where the $\p{N/2}^2$ term contributes a global energy shift which we can neglect, while the $m^2$ term is proportional to $m^2\op{m}=S_\z^2$.
In total, the effective Hamiltonian through second order in perturbation theory is thus
\begin{align}
  H_{\t{eff}}
  = -\f{U}{L}\v S\c\v S - \overline B S_\z + \Omega S_\x + \chi S_\z^2,
\label{eq:H_OAT}
\end{align}
with
\begin{align}
  \chi \equiv \f{\widetilde B^2}{\p{N-1}fU}.
\end{align}
We benchmark the validity of this effective Hamiltonian via exact simulations of the spin [Eqn.~\eqref{eq:SpinModel}] and OAT [Eqn.~\eqref{eq:H_eff}] Hamiltonians in a system of 20 spins, finding that the relative error in maximal squeezing (in dB) of the OAT model is less than 3\% when $\widetilde{B}/U<0.06$\cite{supplementary_material}.

\section{Two-axis twisting, decoherence, and the residual axial field}
\label{sec:derivation_TAT}

The protocol we use to transform one-axis twisting (OAT) into two-axis twisting (TAT) is as previously proposed in Ref.~[\citenum{huang2015twoaxis}]; we provide a summary of this protocol here, in addition to some brief discussion of its implications for decoherence and the residual $\sim\v S\v\c\v S$ and $\sim S_\z$ terms of our OAT protocol.
The TAT protocol begins with the OAT Hamiltonian with a time-dependent transverse field,
\begin{align}
  H = \chi S_\z^2 + \Omega\p{t} S_\x,
  &&
  \Omega\p{t} = \beta\omega \cos\p{\omega t},
  \label{eq:driven_OAT}
\end{align}
where $\beta$ is the modulation index of the driving field and the drive frequency $\omega\gg N\chi$, with $N$ the total number of spins.
Moving into the rotating frame of $\Omega\p{t}S_\x$ subtracts this term from the Hamiltonian, and transforms operators $\O$ as
\begin{align}
  \O \to U\p{t}^\dag \O U\p{t},
\end{align}
where
\begin{align}
  U\p{t} \equiv \exp\sp{-i\int_0^t\d\tau~\Omega\p{\tau}S_\x}
  = \exp\sp{-i\beta\sin\p{\omega t}S_\x}.
\end{align}
In particular, the operators $\tilde S_\pm\equiv-S_\z\pm iS_\y$ (i.e.~the raising and lowering operators in the $x$ basis) transform simply as
\begin{align}
  \tilde S_\pm
  \to U^\dag \tilde S_\pm U
  = e^{\pm i\beta\sin\p{\omega t}} \tilde S_\pm.
\end{align}
For any operator $\O$ and drive frequency $\omega\gg\norm{\O}$, where $\norm{\O}\equiv\max_\psi\sqrt{\braket{\psi|\O^\dag\O|\psi}}$ is the operator norm of $\O$ (i.e.~the magnitude of the largest eigenvalue of $\O$), we can generally make a secular approximation to say
\begin{align}
  e^{\pm im\beta\sin\p{\omega t}} \O
  &= \sum_{n=-\infty}^\infty \J_n\p{\pm m\beta} e^{in\omega t} \O \notag \\
  &\approx \J_0\p{\pm m\beta} \O
  = \J_0\p{m\beta} \O,
\end{align}
where $\J_n$ is the $n$-th order Bessel function of the first kind.
Expanding $S_\z^2=\f14\p{\tilde S_+ + \tilde S_-}^2$, one can thus work out that the effective Hamiltonian in the rotating frame of the drive is
\begin{align}
  H_{\t{eff}}
  \approx \f{\chi}{2} \p{\sp{1+\J_0\p{2\beta}} S_\z^2
  + \sp{1-\J_0\p{2\beta}} S_\y^2}.
\end{align}
Driving with a modulation index $\beta$ for which $J_0\p{2\beta}=\pm1/3$ then gives us the effective two-axis twisting Hamiltonians
\begin{align}
  H_{\t{eff}}^{(+)}
  &= \f{\chi}{3} \p{2 S_\z^2 + S_\y^2}
  \simeq \f{\chi}{3} \p{S_\z^2 - S_\x^2}, \\
  H_{\t{eff}}^{(-)}
  &= \f{\chi}{3} \p{S_\z^2 + 2 S_\y^2}
  \simeq \f{\chi}{3} \p{S_\y^2 - S_\x^2},
\end{align}
where $\simeq$ denotes equality up to the addition of a term proportional to $\v S^2=S_\z^2+S_\x^2+S_\y^2$, which is irrelevant in the absence of coherent coupling between states with different net spin.
In a similar spirit, one can work out that single-spin operators transverse to the $x$-axis transform as
\begin{multline}
  \tilde s_\pm
  \equiv \f12\p{-s_\z\pm i s_\y} \\
  \to U^\dag \tilde s_\pm U
  = e^{\pm i\beta\sin\p{\omega t}} \tilde s_\pm
  \approx \J_0\p{\beta} \tilde s_\pm,
\end{multline}
which implies that shifting into the rotating frame of the time-dependent drive takes
\begin{align}
  s_\x \to s_\x,
  &&
  s_{\y,\z} \to \J_0\p{\beta} s_{\y,\z},
  \label{eq:dec_transformation_xyz}
\end{align}
and
\begin{align}
  s_\pm
  \to \f12\sp{1 \pm \J_0\p{\beta}} s_+
  + \f12\sp{1 \mp \J_0\p{\beta}} s_-.
  \label{eq:dec_transformation_pm}
\end{align}
As the TAT Hamiltonians $H_{\t{eff}}^{(\pm)}$ are realized in a rotating frame, to properly account for decoherence throughout the TAT protocol one must transform jump operators according to Eqns.~\eqref{eq:dec_transformation_xyz}-\eqref{eq:dec_transformation_pm}.

In practice, our protocols realize the OAT Hamiltonian in Eqn.~\eqref{eq:driven_OAT} with additional $\sim\v S\c\v S$ and $\sim S_\z$ terms [see Eqn.~\eqref{eq:H_OAT}].
The effect of the $\sim\v S\c\v S$ term is to generate a relative phase between states with different total spin $S$ (with e.g.~$S=N/2$ within the Dicke manifold).
In the absence of coherent coupling between states with different total spin, therefore, the $\sim\v S\c\v S$ term has no effect on system dynamics.
The $\sim S_z$ term, meanwhile, is important; the magnitude of this term (as measured by the operator norm) is generally comparable to that of the squeezing term $\chi S_\z^2$.
Unlike in the case of OAT, $S_\z$ does not commute with the TAT Hamiltonians, so its effects cannot be eliminated by a single spin-echo $\pi$-pulse $\exp\p{-i\pi S_\x}$ half way through the squeezing protocol.
Nonetheless, we find that for $N=10^2$ ($10^3$) atoms, $\sim5$ (10) $\pi$-pulses in a CPMG (Carr-Purcell-Meiboom-Gill) sequence\cite{carr1954effects, meiboom1958modified} suffice to mitigate the effects of the $S_\z$ term in the TAT protocol (see Appendix \ref{sec:dynamical_decoupling}).
Phase control over these pulses, specifically choices of whether to apply $\exp\p{\pm i\pi S_\x}$ or $\exp\p{\pm i\pi S_\y}$ in any given $\pi$-pulse, can be used to construct XY-$n$ pulse sequences\cite{maudsley1986modified, gullion1990new} that are robust to pulse errors.

\section{Dynamical decoupling in the TAT protocol}
\label{sec:dynamical_decoupling}
\setcounter{figure}{0}

The effective Hamiltonian resulting from a perturbative treatment of SOC is (see Appendix \ref{sec:derivation_OAT})
\begin{align}
  H_{\t{eff}}
  = -\f{U}{L}\v S\c\v S - \overline B S_\z + \Omega S_\x + \chi S_\z^2,
  \label{eq:H_eff_DD}
\end{align}
where $U$ is a two-atom on-site interaction strength; $L$ is the number of lattice sites; $\overline B\equiv \sum_n B_n/N$ is a residual axial field determined by the occupied quai-momentum modes $\set{n}$ (with $\abs{\set{n}}=N$ atoms total); $\Omega$ is the magnitude of a driving field; and $\chi$ is an effective OAT squeezing strength.
The effect of the $\sim\v S\c\v S$ term is to generate a relative phase between states with different total spin $S$ (where $S=N/2$ within the Dicke manifold).
In the absence of coherent coupling between states with different total spin, therefore, the $\sim\v S\c\v S$ term has no effect on system dynamics, and we are safe to neglect it entirely.

In the parameter regimes relevant to our discussions in the main text, the operator norms of $\overline{B}\hat S_\z$ and $\chi \hat S_\z^2$ in Eqn.~\eqref{eq:H_eff_DD} will typically be comparable in magnitude.
The OAT protocol sets $\Omega=0$, and eliminates the effect of $\overline{B}\hat S_\z$ with a spin-echo $\pi$-pulse $\exp\p{-i\pi\hat S_\x}$ applied half way through the squeezing protocol.
The TAT protocol, meanwhile, effectively takes $\chi\hat S_\z^2+\Omega \hat S_\x\to\hat H_{\t{TAT}}^{(\pm)}$ (as defined in Appendix \ref{sec:derivation_TAT}) and $\overline{B}\hat S_\z\to\J_0\p{\beta_\pm}\overline{B}\hat S_\z$, where $\J_0$ is the zero-order Bessel function of the first kind and $\beta_\pm$ is the modulation index of the amplitude-modulated driving field $\Omega$, satisfying $\J_0\p{2\beta_\pm}=\pm1/3$.
Unlike in the case of OAT, $\hat S_\z$ does not commute with the TAT Hamiltonian, so its effect cannot be eliminated with a spin-echo.
Nonetheless, this term can be eliminated with a dynamical decoupling pulse sequence that periodically inverts the sign of $\hat S_\z$ while preserving $\hat H_{\t{TAT}}^{(\pm)}$.

\begin{figure*}
\centering
\includegraphics{pulsed_squeezing.pdf}
\caption{{\bf Optimal squeezing as a function of $\pi$-pulses} applied prior to the optimal TAT squeezing time in a CPMG sequence with ({\bf a)} $N=100$ and ({\bf b}) $N=1000$ atoms.
Results are shown for OAT, TAT, and TAT$_{\pm,\z}$, where TAT$_{\pm,\z}$ denotes squeezing via the Hamiltonian $\hat H_{\t{TAT}}^{(\pm,\z)}\equiv \hat H_{\t{TAT}}^{(\pm)}-\J_0\p{\beta_\pm}\bk{\overline B}^{\t{rms}}_f \hat S_\z$.
Details about experimental parameters for these simulations are provided in the text.
}
\label{fig:pulsed_squeezing}
\end{figure*}

Fig.~\ref{fig:pulsed_squeezing} shows the maximal squeezing generated by $N=10^2$ and $10^3$ atoms via OAT, TAT, and TAT in the presence of the mean field $\J_0\p{\beta_\pm}\overline B\hat S_\z$ as a function of the number of $\pi$-pulses performed prior to the optimal TAT squeezing time.
These pulses are applied in a CPMG sequence $\p{\tau_n/2-\pi_\x-\tau_n/2}^n$, where $\tau_n/2$ denotes Hamiltonian evolution for a time $\tau_n/2$, $\pi_\x$ denotes the application of an instantaneous $\pi$-pulse $\exp\p{-i\pi\hat S_\x}$, and $n$ is the number of pulses, such that the optimal TAT squeezing time is $t_{\t{opt}}^{\t{TAT}}=\p{\tau_n}^n$.
The label TAT$_{\pm,\z}$ in Fig.~\ref{fig:pulsed_squeezing} denotes squeezing through the Hamiltonian $\hat H_{\t{TAT}}^{(\pm,\z)}\equiv \hat H_{\t{TAT}}^{(\pm)}-\J_0\p{\beta_\pm}\bk{\overline B}^{\t{rms}}_f\hat S_\z$, where $\bk{\overline B}^{\t{rms}}_f$ is the root-mean-square average of $\overline B$ over choices of occupied spacial modes $\set{n}$ at fixed filling $f$ of all spatial modes in the lowest Bloch band of a periodic 2D lattice.
While the modulation index $\beta_+$ is uniquely defined by $\J_0\p{2\beta_+}=1/3$, there are two choices of $\beta_-$ for which $\J_0\p{2\beta_-}=-1/3$; we use that which minimizes $\abs{\J_0\p{\beta_-}}$.
Fig.~\ref{fig:pulsed_squeezing} assumes an SOC angle $\phi=\pi/50$ (although results are independent of $\phi$ for $\phi\ll1$), a reduced field variance $\widetilde{B}/U=0.05$, and a filling $f=5/6$.
Note that as the filling $f\to1$, the residual axial field vanishes ($\overline{B}\to0$), so TAT$_{\pm,\z}\to$ TAT.

\section{Clock interrogation after squeezing}
\label{sec:clock_interrogation}

The protocols in our work concern the preparation of spin-squeezed states in an optical lattice clock.
Here, we discuss the use of these states in a follow-up clock interrogation protocol.
For simplicity, we restrict our discussion to the case of squeezing in 1D, as in Section \ref{sec:theory} of the main text, with the understanding that a generalization of this discussion to higher dimensions is straightforward.

A spin-squeezed state is generated by interactions and SOC that are generally undesirable during the clock interrogation protocol.
In the parameter regimes considered in our work, interactions alone have no effect on clock interrogation: absent of coherence between states with different net spin $S$, collective $\v S\c\v S$ interactions only generate unobservable global phases within each fixed-$S$ sector of Hilbert space.
The remaining task to allow clock interrogation after spin squeezing is therefore to turn off SOC, which inhomogeneously detunes atomic transition frequencies by an amount $B_q$ that depends on the quasi-momentum $q$ of an atom.
The SOC-induced axial fields $B_q\sim J\sin(\phi/2)$ depend on two tunable parameters: the tunneling rate $J$ and the SOC angle $\phi$.
The simplest way to turn off SOC is thus to increase the lattice depth prior to clock interrogation, taking $J\sim B_q\to0$.
Increasing the lattice depth to turn off SOC is compatible with the current clock interrogation sequence, but is incompatible with ongoing efforts to mitigate light scattering from the lattice beams (currently the primary source of decoherence in the clock; see Appendix \ref{sec:decoherence}) by use of a shallow lattice\cite{hutson2019engineering}.
We therefore devote the rest of this section to discussing strategies for turning off SOC that are compatible with the use of shallow lattices.

If we cannot take the tunneling rate $J\to0$, our remaining control parameter for turning off SOC is the SOC angle $\phi=k_{\t{L}}a$, where $k_{\t{L}}$ is the projection of the clock laser wavenumber onto the lattice axis and $a$ is the lattice spacing.
The squeezing protocol needs a clock laser with a small but nonzero SOC angle $\phi\ll1$, while the clock interrogation protocol requires a clock laser with $\phi=0$.
Simply using one clock laser with $\phi\ne0$ for the squeezing protocol and another clock laser with $\phi=0$ for the clock interrogation protocol, however, does not resolve this discrepancy, because (i) both the squeezing and clock interrogation protocols need a clock laser that is homogeneous -- i.e.~with Hamiltonian proportional to $S_\x$, (ii) lasers with different SOC angles are homogeneous in different gauges, and (iii) spin squeezing is a gauge-dependent quantity.
More precisely, a state that is squeezed with respect to spin operators that are homogeneous (i.e.~of the form $S_\x,S_\y,S_\z$) in a particular gauge is not necessarily squeezed with respect to spin operators that are homogeneous in a different gauge.
To address this problem, we will fix a gauge in which the Fermi-Hubbard Hamiltonian exhibits SOC, as in Eqn.~\eqref{eq:FermiHubbard}, and construct a simple pulse sequence that transforms the $\phi=0$ clock laser into homogenous form in our working gauge.

Defining on-site spin operators
\begin{align}
  s_\z^{(j)} &\equiv \f12 \p{ c_{j,\up}^\dag c_{j,\up}
    - c_{j,\dn}^\dag c_{j,\dn}} \\
  s_\x^{(j)} &\equiv \f12 \p{c_{j,\up}^\dag c_{j,\dn}
    + c_{j,\dn}^\dag c_{j,\up}} \\
  s_\y^{(j)} &\equiv \f{i}{2} \p{c_{j,\dn}^\dag c_{j,\up}
    - c_{j,\up}^\dag c_{j,\dn}},
\end{align}
we observe that gauge transformation $\hat c_{j,\up}\to e^{i\theta j}\hat c_{j,\up}$ takes (up to an overall phase)
\begin{align}
  S_\x = \sum_j s_\x^{(j)}
  \to \sum_j \p{\cos\p{\theta j}  s_\x^{(j)}
    + \sin\p{\theta j} s_\y^{(j)}}
  \equiv S_\x^{(\theta)},
\end{align}
This transformation thus rotates local spin operators by a site-dependent angle $\theta j$, generally taking $S_{\x,\y}^{(\alpha)}\to S_{\x,\y}^{(\alpha+\theta)}$ (with $S_\y^{(\alpha)}$ defined similarly to $S_\x^{(\alpha)}$) and leaving $S_\z$ unaffected (taking $S_\z\to S_\z$).
A clock laser with SOC angle $\phi$ induces a Hamiltonian proportional to $S_\x^{(-\phi)}$ ($S_\x^{(-\phi+\theta)}$) before (after) a gauge transformation by $\theta$, which is why transforming with $\theta=\phi$ made the SOC-inducing clock laser homogenous in the main text.

To ``undo'' the effect of our gauge transformation on spin operators, we can construct the unitary ``gauge-switching'' operator
\begin{align}
  G_\theta = \prod_j \exp\p{-i \theta j s_\z^{(j)}}
  \label{eq:gauge_rotation}
\end{align}
for which
\begin{align}
  G_\theta^\dag S_{\x,\y}^{(\alpha+\theta)} G_\theta = S_{\x,\y}^{(\alpha)}.
\end{align}
We then decompose any rotation into a product of two reflections:
\begin{align}
  \exp\p{-i\alpha s_\z^{(j)}}
  = \exp\p{-i\pi s_\x^{(j)}} \exp\p{-i\pi s_{\alpha/2}^{(j)}},
  \label{eq:rotation_reflections}
\end{align}
where $s_{\alpha/2}^{(j)} \equiv \cos\p{\alpha/2}s_\x^{(j)} + \sin\p{\alpha/2}s_\y^{(j)}$.
This decomposition implies
\begin{align}
  G_\theta = \exp\p{-i\pi S_\x^{(\theta/2)}} \exp\p{-i\pi S_\x^{(\theta)}}.
  \label{eq:gauge_decomposition}
\end{align}
The unitary in Eqns.~\eqref{eq:gauge_rotation} and \eqref{eq:gauge_decomposition} can therefore be implemented by a sequence of two $\pi$-pulses: one $\pi$-pulse by a clock laser with SOC angle $\phi=0$ (with Hamiltonian $\sim S_\x^{(\theta)}$ in the gauged frame), and a second $\pi$-pulse by a clock laser with SOC angle $\phi=\theta/2$ (with Hamiltonian $\sim S_\x^{(\theta/2)}$ in the gauged frame).
This unitary can be used to reconcile the incommensurate SOC requirements of the squeezing and clock interrogation protocols.

Starting with a spin-down-ploarized initial state
\begin{align}
  \ket{-\t{Z}}\equiv\p{\prod_j c_{j,\dn}^\dag}\ket{\t{vacuum}},
\end{align}
our OAT protocol essentially prepares the squeezed state
\begin{align}
  \ket{\xi_{\t{OAT}}}
  \equiv e^{-iH_{\t{FH}}^\theta t}
  e^{-i\p{\pi/2} S_\x} \ket{-\t{Z}},
\end{align}
for some free evolution time $t$ and a Fermi-Hubbard Hamiltonian $H_{\t{FH}}^\theta$ exhibiting SOC with angle $\theta\ll1$, where $\theta=\phi$ in Eqn~\eqref{eq:FermiHubbard}.
The follow-up clock interrogation protocol needs to rotate this squeezed state, most generally by some unitary $\exp\p{-i\v\alpha\c\v S}$, and then measure collective spin observables of the form (for example)
\begin{align}
  \bk{S_\x}_{\t{OAT}}^{\v\alpha}
  \equiv \bk{\xi_{\t{OAT}}|e^{i\v\alpha\c\v S}
    S_\x e^{-i\v\alpha\c\v S}|\xi_{\t{OAT}}}.
\end{align}
In a clock interrogation protocol with a shallow lattice, the rotation $\exp\p{-i\v\alpha\c\v S}$ and measurement of $S_\x$ must be performed with an SOC-free ($\phi=0$) clock laser, which can effectively only access the gauge-transformed operators $S_{\x,\y}^{(\theta)}$.
We therefore expand
\begin{align}
  \bk{S_\x}_{\t{OAT}}^{\v\alpha}
  &= \bk{\xi_{\t{OAT}}| G_\theta^\dag e^{i\v\alpha\c\v S^{(\theta)}}
    S_\x^{(\theta)} e^{-i\v\alpha\c\v S^{(\theta)}}
    G_\theta |\xi_{\t{OAT}}} \notag \\
  &= \bk{\xi_{\t{OAT}}^\theta| e^{i\v\alpha\c\v S^{(\theta)}}
    S_\x^{(\theta)} e^{-i\v\alpha\c\v S^{(\theta)}} |\xi_{\t{OAT}}^\theta},
\end{align}
where $\v S^{(\theta)}\equiv G_\theta \v S G_\theta^\dag$ and
\begin{align}
  \ket{\xi_{\t{OAT}}^\theta} \equiv G_\theta \ket{\xi_{\t{OAT}}}.
\end{align}
We are thus free to use the SOC-free clock laser during clock interrogation if we append the operation $G_\theta$ to our squeezing protocol.

The last subtlety of clock interrogation and ``gauge-switching'' concerns the squeezing protocol itself.
Experimentally implementing $G_\theta$ to enable clock interrogation requires clock lasers with SOC angles $\phi=0$ and $\phi=\theta/2$, neither of which is homogenous in our working gauge.
Our squeezing protocols, however, need a homogenous clock laser for (i) an initial $\pi/2$ rotation of $\ket{-\t{Z}}$, (ii) spin-echo pulses throughout the squeezing protocol, and (iii) the drive used to transform OAT into TAT.
Any clock laser pulses in the squeezing protocols can still be implemented using $G_\theta$ and the SOC-free ($\phi=0$) clock laser via the same trick used for clock interrogation, e.g.~by decomposing
\begin{align}
  \ket{\xi_{\t{OAT}}^\theta}
  &= G_\theta e^{-iH_{\t{FH}}^\theta t}
  G_\theta^\dag e^{-i\p{\pi/2} S_\x^{(\theta)}} G_\theta \ket{-\t{Z}},
  \notag \\
  &\simeq G_\theta e^{-iH_{\t{FH}}^\theta t}
  G_\theta^\dag e^{-i\p{\pi/2} S_\x^{(\theta)}} \ket{-\t{Z}},
\end{align}
where $\simeq$ denotes equality up to a global phase, and we used the fact that the initial state $\ket{-\t{Z}}$ is an eigenstate of $G_\theta$.
Applying a continuous drive, meanwhile, seemingly requires an extra clock laser with SOC angle $\phi=\theta$.
In practice, it may be prohibitively difficult to properly calibrate the relative angles of three clock lasers.
% note to authors: the use of a third laser is also incompatible with a 3D accordion lattice, but I (MP) decided that this point would require too much extra discussion to explain here
In this case, converting OAT into TAT would need to be done with a pulsed drive protocol, as e.g.~in Ref.[\citenum{liu2011spin}].

\section{Decoherence in the 3D $^{87}$Sr optical lattice clock}
\label{sec:decoherence}

Currently, light scattering from lattice beams in the 3D $^{87}$Sr optical lattice clock induces decoherence on a time scale of $\sim$10 seconds\cite{goban2018emergence, hutson2019engineering}.
This decoherence acts identically on all atoms in an uncorrelated manner, and can be understood by considering the density operator $\rho$ for a single atom, with effective spin states $\dn$ and $\up$ respectively corresponding to the ${}^1\t{S}_0$ and ${}^3\t{P}_0$ electronic states.
Empirically, the effect of decoherence after a time $t$ within the $\set{\dn,\up}$ subspace of a single atom is to take $\rho\to\rho\p{t}$ with $\rho\p{0}\equiv\rho$ and
\begin{align}
  \rho\p{t} \coloneqq
  \begin{pmatrix}
  \rho_{\up\up} e^{-\Gamma_{\up\up}t} &&
  \rho_{\up\dn} e^{-\Gamma_{\up\dn}t} \\
  \rho_{\up\dn}^* e^{-\Gamma_{\up\dn}t} &&
  \rho_{\dn\dn} + \p{1-e^{-\Gamma_{\up\up}t}} \rho_{\up\up}
  \end{pmatrix},
\end{align}
where $\Gamma_{\up\up}\approx\Gamma_{\up\dn}\approx\Gamma=0.1~\t{sec}^{-1}$ are respectively decay rates for $\rho_{\up\up}$ and $\rho_{\up\dn}$.
This form of decoherence can be effectively modeled by decay and dephasing of individual spins (respectively denoted $\Gamma_{\t{ud}}$ and $\Gamma_{\t{el}}$ in Ref.~[\citenum{foss-feig2013nonequilibrium}]) at rates $\Gamma$.
In the language of the section that follows, we would say that this decoherence is captured by the sets of jump operators $\J_-\equiv\set{s_-^{(j)}}$ and $\J_\z\equiv\set{s_\z^{(j)}}$ with corresponding decoherence rates $\gamma_-=\gamma_\z=\Gamma$.

\section{Solving Heisenberg equations of motion for collective spin systems}
\label{sec:collective_simulation}

In order to compute squeezing of a collective spin system, we need to compute expectation values of collective-spin operators.
Choosing the basis $\set{\S_{\v m}}$ for these operators, where $\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$ with
$\v m\equiv\p{m_+,m_\z,m_-}\in\mathbb{N}_0^3$, we can expand all
collective-spin Hamiltonians in the form
\begin{align}
  H = \sum_{\v m} h_{\v m} \S_{\v m}.
  \label{eq:H_general}
\end{align}
The evolution of a general operator $\S_{\v m}$ under a Hamiltonian of the form in Eqn.~\eqref{eq:H_general} is then given by
\begin{align}
  \f{d}{dt} \S_{\v m}
  &= i \sum_{\v n} h_{\v n}
  \sp{\S_{\v n}, \S_{\v m}}_-
  + \sum_\J \gamma_\J \mathcal D\p{\J; \S_{\v m}} \notag \\
  &\equiv \sum_{\v n} T_{\v m\v n} \S_{\v n},
\end{align}
where $\sp{X,Y}_\pm\equiv XY\pm YX$; $\J$ is a set of jump operators with corresponding decoherence rate $\gamma_\J$; the decoherence operator $\mathcal D$ is defined by
\begin{align}
  \mathcal{D}\p{\J;\O}
  \equiv \sum_{J\in\J}\p{J^\dag \O J - \f12\sp{J^\dag J,\O}_+};
\end{align}
and $T_{\v m\v n}$ is a matrix element of the time derivative operator $T\equiv d/dt$.
These matrix elements can be calculated analytically using product and commutation rules for collective spin operators.
We can then expand Heisenberg operators in a Taylor series about $t=0$ to write
\begin{align}
  \bk{\S_{\v m}}
  &= \sum_{k\ge0} \f{t^k}{k!} \bk{\f{d^k}{dt^k} \S_{\v m}}_{t=0} \notag \\
  &= \sum_{k\ge0} \f{t^k}{k!}
  \sum_{\v n} T_{\v m\v n;k} \bk{\S_{\v n}}_{t=0},
  \label{eq:time_series}
\end{align}
where $T_{\v m\v n;k}\equiv\sp{T^k}_{\v m\v n}$ are matrix elements of the $k$-th time derivative.
Expectation values of collective spin operators can thus be computed via the expansion in Eqn.~\eqref{eq:time_series}, which at short times can be truncated at some finite order beyond which all terms have negligible contribution to $\bk{\S_{\v m}}$.

\bibliography{main.bib}

\end{document}

### 2019-05-21 soc_squeezing/squeezing_paper/supplementary.tex
added soc squeezing paper in the middle of review last thing worked on: Appendix D in the main text

\documentclass[prx,superscriptaddress,12pt]{revtex4-2}

%%% linking references
\usepackage{hyperref}
\hypersetup{
  colorlinks=true,
  linkcolor=blue,
  urlcolor=cyan,
}

%%% physics and math packages
\usepackage{physics} % general physics package
\usepackage{amssymb} % math fonts and symbols
\usepackage{bm} % for making math symbols bold
\usepackage{braket} % for nice brackets (e.g. |X>)
\usepackage{mathtools} % for \coloneqq
\usepackage{graphicx} % for figures
\graphicspath{{./figures/}} % set path for all figures

%%% shorthands to use inside math environments
\renewcommand{\t}{\text} % text in math mode
\newcommand{\f}[2]{\dfrac{#1}{#2}} % shorthand for fractions
\newcommand{\p}[1]{\left(#1\right)} % parenthesis
\renewcommand{\sp}[1]{\left[#1\right]} % square parenthesis
\renewcommand{\set}[1]{\left\{#1\right\}} % square parenthesis
\renewcommand{\c}{\cdot} % inner product
\renewcommand{\d}{\text{d}} % integration measure, e.g. d^3 x
\renewcommand{\v}{\bm} % bold symbols (e.g. for vectors)
\newcommand{\uv}[1]{\hat{\bm #1}} % unit vectors, or vector operators
\newcommand{\bk}{\Braket}
\renewcommand{\ket}{\Ket}
\renewcommand{\bra}{\Bra}

\newcommand{\B}{\mathcal{B}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\G}{\mathcal{G}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\J}{\mathcal{J}}
\renewcommand{\L}{\mathcal{L}}
\renewcommand{\O}{\mathcal{O}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\Q}{\mathcal{Q}}
\renewcommand{\S}{\mathcal{S}}

\newcommand{\z}{\text{z}}
\newcommand{\x}{\text{x}}
\newcommand{\y}{\text{y}}

\newcommand{\up}{\uparrow}
\newcommand{\dn}{\downarrow}

%%% text for affiliations and contributions
\newcommand{\JILA}{JILA, National Institute of Standards and Technology and Department of Physics, University of Colorado, Boulder, CO, 80309, USA}
\newcommand{\CTQM}{Center for Theory of Quantum Matter, University of Colorado, Boulder, CO, 80309, USA}
\newcommand{\contrib}{\thanks{Authors P.H.~and M.A.P.~contributed equally to this work.}}

%%% add "S" prefix to equation and figure numbers
\renewcommand{\thesection}{S\arabic{section}}
\renewcommand{\theequation}{S\arabic{equation}}
\renewcommand{\thefigure}{S\arabic{figure}}

%%% for referencing files in the main text
\usepackage{xr}
\externaldocument[M-]{main}

\begin{document}

\title{Supplementary Material for ``Engineering spin squeezing in a 3D optical lattice with interacting spin-orbit-coupled fermions''}

\author{P.~He} \contrib
\author{M.A.~Perlin} \contrib
\author{S.R.~Muleady}
\author{R.J.~Lewis-Swan}
\affiliation{\JILA}
\affiliation{\CTQM}
\author{R.B.~Hutson}
\author{J.~Ye}
\affiliation{\JILA}
\author{A.M.~Rey}
\affiliation{\JILA}
\affiliation{\CTQM}

\maketitle

\section{Benchmarking}

Here we provide additional information about our benchmarking of the one-axis twisting model against the spin model in the main text.
This benchmarking was performed via exact simulations of a 20-spin system.
Fig.~\ref{fig:relative_error} shows the relative error in maximal squeezing of the OAT model (measured against the spin model) as a function of the reduced field variance $\widetilde{B}/U$.
Here squeezing is measured in decibels (dB) by $-10\log_{10}\xi^2$ for the squeezing parameter $\xi^2\equiv\min_\theta\braket{\t{var}(\hat S^\perp_\theta)}\times N/\abs*{\braket{\hat{\v S}}}^2$ as defined in the the main text.
The relative error in maximal squeezing (in dB) by the OAT model is less than 3\% when $\widetilde{B}/U<0.06$.

In principle, spin-changing decoherence compromises the validity of the OAT model, as its perturbative derivation in Appendix \ref{M-sec:derivation_OAT} of the main text relies on spin population remaining primarily within the Dicke manifold.
This assumption breaks down in the presence of, for example, spontaneous emission, which transfers population outside of the Dicke manifold.
Nonetheless, we find decent agreement between the OAT and spin models when decoherence is sufficiently weak (see Fig.~\ref{fig:benchmarking_decay}).

\begin{figure*}
\centering
\includegraphics[width=0.5\textwidth]{relative_error.pdf}
\caption{{\bf Relative error} between maximal squeezing (measured in dB) obtained by the OAT [Eqn.~(3)] and spin [Eqn.~(2)] models of the main text in a system of 20 particles.
The OAT model correctly captures the maximal squeezing (in dB) of the spin model to within 3\% (marked by the horizontal reference line) within the gap-protected regime $\widetilde{B}/U<0.06$.
}
\label{fig:relative_error}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=1.0\textwidth]{benchmarking_decay.pdf}
\caption{{\bf Comparison between the OAT and the spin model in the presence of decoherence.}
({\bf a}) The difference between the maximal squeezing (measured in dB) obtained by the OAT [Eqn.~(3)] and spin [Eqn.~(2)] models increases with the particle number $N$ and the single-particle spontaneous emission rate $\gamma$.
This disagreement is attributed in part to the fact that spontaneous emission transfers population of the collective spin state outside of the Dicke manifold, violating an assumption of the OAT model; see panel ({\bf b}).
The rate of population transfer outside of the Dicke manifold increases with both particle number and spontaneous emission rate.
(Parameters for simulations in this figure: $U=1000$ Hz, $J=200$ Hz, and $\phi=\pi/20$).
}
\label{fig:benchmarking_decay}
\end{figure*}

\section{Time-series of squeezing via OAT and TAT}

Figure \ref{fig:squeezing_example} shows an example of squeezing over time via OAT and TAT, both with and without decoherence via decay and dephasing of individual spins.
The OAT model initially generates squeezing faster than the TAT model, but the squeezing generation rate of OAT (measured in dB per second) falls off with time.
The squeezing generation rate for TAT, meanwhile, remains approximately constant (in the absence of decoherence) until squeezing via TAT surpasses that of OAT.
In the absence of decoherence, OAT achieves a maximal amount of squeezing that scales as $\xi^2\sim N^{-2/3}$, while TAT achieves Heisenberg-limited squeezing with $\xi^2\sim N^{-1}$.
Note that our method for computing squeezing via TAT in the presence of decoherence (described in Appendix \ref{M-sec:collective_simulation} of the main text) is not capable of computing squeezing for the full range of times shown in Fig.~\ref{fig:squeezing_example}; the corresponding time-series data in this figure is therefore shown up to the point at which this method breaks down.

\begin{figure*}
\centering
\includegraphics{squeezing_example_L40.pdf}
\includegraphics{squeezing_example_L100.pdf}
\caption{{\bf Squeezing via OAT and TAT} in a 2D section of the 3D $^{87}$Sr optical lattice clock, shown for ({\bf a}) $\ell=40$ and ({\bf b}) $\ell=100$ sites per axis (with $N=\ell^2$ atoms total), and a lattice depth of $V_0=4~E_{\t{R}}$, where $E_{\t{R}}$ is the atomic lattice recoil energy.
Atoms are confined along the direction transverse to the 2D layer by a lattice of depth 60 $E_{\t{R}}$.
Squeezing over time is shown for OAT (blue) and TAT (green), both with (solid lines) and without (dashed lines) decoherence via uncorrelated decay and dephasing of individual spins at rates of $0.1~\t{sec}^{-1}$ (see Appendix \ref{M-sec:decoherence} of the main text).
}
\label{fig:squeezing_example}
\end{figure*}

\end{document}

### 2019-05-21 soc_squeezing/squeezing_paper/main.tex
rearranged Appendix D in squeezing paper

@@
 \begin{align}
-  \hat H_{\text{FH}}/\hbar
+  \hat H_{\text{FH}}^{(\phi)}/\hbar
   &= -J \sum_j \p{e^{i\phi} \hat c_{j,\up}^\dag \hat c_{j+1,\up}
@@
 \begin{align}
-  \hat H_{\t{FH,single}}/\hbar
+  \hat H_{\t{FH,single}}^{(\phi)}/\hbar
   = -2J\sum_q\sp{\cos(qa + \phi) \hat n_{q,\up}
@@
 In the weak SOC limit (i.e.~$B_q\ll fU$), however, the interaction energy gap suppresses population transfer between states with different total spin $S$ [see Fig.~\ref{fig:protocol_schematic}(d)].
-%In this weak SOC limit (i.e.~of all $B_q/fU\ll1$), we can treat the axial field perturbatively and perform a Schrieffer-Wolff transformation to determine its effective action on states restricted to the Dicke manifold (see Appendix \ref{sec:derivation_OAT}).
 In this regime, the virtual occupation of states outside the Dicke manifold can be accounted for perturbatively.
@@
 can be determined experimentally from global spin measurements.
-Here $\braket{\hat{\v S}}$ is the mean collective-spin vector and $\braket{\t{var}(\hat S^\perp_\theta)}$ is the variance of spin measurements along an axis orthogonal to $\braket{\hat{\v S}}$, parameterized by the angle $\theta\in[0,2\pi)$.
+Here $\braket{\hat{\v S}}$ is the mean collective spin vector and $\braket{\t{var}(\hat S^\perp_\theta)}$ is the variance of spin measurements along an axis orthogonal to $\braket{\hat{\v S}}$, parameterized by the angle $\theta\in[0,2\pi)$.
 The above protocol concerns only the preparation of a spin-squeezed state, which would then be used as an input state for a follow-up clock interrogation protocol without SOC.
@@
 More precisely, a state that is squeezed with respect to spin operators that are homogeneous (i.e.~of the form $S_\x,S_\y,S_\z$) in a particular gauge is not necessarily squeezed with respect to spin operators that are homogeneous in a different gauge.
-To address this problem, we will fix a gauge in which the Fermi-Hubbard Hamiltonian exhibits SOC, as in Eqn.~\eqref{eq:FermiHubbard}, and construct a simple pulse sequence that transforms the $\phi=0$ clock laser into homogenous form in our working gauge.
+To resolve the fact that our squeezing protocol prepares a state that is squeezed in the ``wrong gauge'', we will work explicitly in the ``lab gauge'' of the clock interrogation protocol, in which the Fermi-Hubbard Hamiltonian is SOC-free and the $\phi=0$ clock laser is homogenous.
+We will then construct a simple pulse sequence that transforms the inhomogeneous spin operators accessible by the $\phi\ne0$ laser into homogeneous form in our working gauge.
 
+Starting with a spin-down-ploarized initial state
+\begin{align}
+  \ket{-\t{Z}}\equiv\p{\prod_j c_{j,\dn}^\dag}\ket{\t{vacuum}},
+\end{align}
+our OAT protocol essentially prepares the state
+\begin{align}
+  \ket{\xi_{\t{OAT}}^{(\theta)}} = e^{-i H_{\t{FH}}^{(0)}t}
+  e^{-i\p{\pi/2}S_\x^{(\theta)}} \ket{-\t{Z}},
+\end{align}
+where $H_{\t{FH}}^{(0)}$ is the Fermi-Hubbard Hamiltonian in Eqn.~\eqref{eq:FermiHubbard} without SOC; $t$ is some free evolution time; and $S_\x^{(\theta)}$ is the ``rotated'' spin-$x$-like generator induced by a clock laser with SOC angle $\phi=\theta$, namely
+\begin{align}
+  S_\x^{(\theta)}
+  = \f12 \sum_j e^{i\theta j} c_{j,\up}^\dag c_{j,\dn} + \t{h.c.}.
+\end{align}
 Defining on-site spin operators
@@
 \end{align}
-we observe that gauge transformation $\hat c_{j,\up}\to e^{i\theta j}\hat c_{j,\up}$ takes (up to an overall phase)
+we can identify the rotated collective spin operators
 \begin{align}
-  S_\x = \sum_j s_\x^{(j)}
-  \to \sum_j \p{\cos\p{\theta j}  s_\x^{(j)}
-    + \sin\p{\theta j} s_\y^{(j)}}
-  \equiv S_\x^{(\theta)},
+  S_\x^{(\theta)} &\equiv \sum_j \p{\cos\p{\theta j} s_\x^{(j)}
+  + \sin\p{\theta j} s_\y^{(j)}},
+  \label{eq:S_x_theta} \\
+  S_\y^{(\theta)} &\equiv \sum_j \p{\cos\p{\theta j} s_\y^{(j)}
+  - \sin\p{\theta j} s_\x^{(j)}}.
+  \label{eq:S_y_theta}
 \end{align}
-This transformation thus rotates local spin operators by a site-dependent angle $\theta j$, generally taking $S_{\x,\y}^{(\alpha)}\to S_{\x,\y}^{(\alpha+\theta)}$ (with $S_\y^{(\alpha)}$ defined similarly to $S_\x^{(\alpha)}$) and leaving $S_\z$ unaffected (taking $S_\z\to S_\z$).
-A clock laser with SOC angle $\phi$ induces a Hamiltonian proportional to $S_\x^{(-\phi)}$ ($S_\x^{(-\phi+\theta)}$) before (after) a gauge transformation by $\theta$, which is why transforming with $\theta=\phi$ made the SOC-inducing clock laser homogenous in the main text.
+The state $\ket{\xi_{\t{OAT}}^{(\theta)}}$ is squeezed with respect to components of the rotated collective spin vector $\v S_\theta \equiv \p{S_\x^{(\theta)},S_\y^{(\theta)},S_\z}$.
+To take advantage of the squeezing in $\ket{\xi_{\t{OAT}}^{(\theta)}}$, therefore, the clock interrogation protocol effectively needs to rotate this state by some unitary $\exp\p{-i\v\eta\c\v S_\theta}$, and then extract information about the rotation vector $\v\eta$ from collective spin observables of the form
+\begin{align}
+  \bk{\O_\theta}_{\t{OAT}}^{\v\eta}
+  \equiv \bk{\xi_{\t{OAT}}^{(\theta)}| e^{i\v\eta\c\v S_\theta}
+    \O_\theta e^{-i\v\eta\c\v S_\theta} |\xi_{\t{OAT}}^{(\theta)}},
+\end{align}
+where $\O_\theta$ is some product of the rotated collective spin operators in $\v S_\theta$, e.g.~$S_\x^{(\theta)}$ or $S_\x^{(\theta)} S_\y^{(\theta)}$.
+In order to turn off SOC in a shallow lattice, however, the clock interrogation protocol is restricted to performing rotations of the form $\exp\p{-i\v\eta\c\v S_0}$ and measuring homogeneous operators $\O_0$.
+We therefore seek a ``gauge-switching'' operation $G_\theta$ that maps homogenous operators $\O_0$ onto rotated operators $\O_\theta$ via $G_\theta^\dag \O_0 G_\theta = \O_\theta$.
+Equipped with $G_\theta$, we could decompose
+\begin{align}
+  \bk{\O_\theta}_{\t{OAT}}^{\v\eta}
+  &= \bk{\xi_{\t{OAT}}^{(\theta)}| G_\theta^\dag e^{i\v\eta\c\v S_0}
+    \O_0 e^{-i\v\eta\c\v S_0} G_\theta |\xi_{\t{OAT}}^{(\theta)}}
+  \notag \\
+  &= \bk{\tilde \xi_{\t{OAT}}^{(\theta)}| e^{i\v\eta\c\v S_0}
+    \O_0 e^{-i\v\eta\c\v S_0} |\tilde \xi_{\t{OAT}}^{(\theta)}}
+\end{align}
+for a transformed state
+\begin{align}
+  \ket{\tilde \xi_{\t{OAT}}^{(\theta)}}
+  \equiv G_\theta \ket{\xi_{\t{OAT}}^{(\theta)}}
+\end{align}
+that is now squeezed with respect to the homogeneous collective spin operators in $\v S_0$, accessible with the SOC-free ($\phi=0$) clock laser used during clock interrogation.
 
-To ``undo'' the effect of our gauge transformation on spin operators, we can construct the unitary ``gauge-switching'' operator
+Given the definitions of the rotated spin operators $S_\x^{(\theta)},S_\x^{(\theta)}$ in Eqns.~\eqref{eq:S_x_theta}, \eqref{eq:S_y_theta}, a suitable candidate for the gauge-switching operator $G_\theta$ is the site-dependent rotation
 \begin{align}
-  G_\theta = \prod_j \exp\p{-i \theta j s_\z^{(j)}}
-  \label{eq:gauge_rotation}
+  G_\theta = \prod_j \exp\p{i\theta j s_\z^{(j)}}.
 \end{align}
-for which
+To implement $G_\theta$ with ``global'' (i.e.~non-site-selective) experimental controls, we decompose any rotation into a product of two reflections:
 \begin{align}
-  G_\theta^\dag S_{\x,\y}^{(\alpha+\theta)} G_\theta = S_{\x,\y}^{(\alpha)}.
+  \exp\p{i2\alpha s_\z^{(j)}}
+  \simeq \exp\p{i\pi s_\x^{(j)}} \exp\p{i\pi s_\alpha^{(j)}}
 \end{align}
-We then decompose any rotation into a product of two reflections:
-\begin{align}
-  \exp\p{-i\alpha s_\z^{(j)}}
-  = \exp\p{-i\pi s_\x^{(j)}} \exp\p{-i\pi s_{\alpha/2}^{(j)}},
-  \label{eq:rotation_reflections}
-\end{align}
-where $s_{\alpha/2}^{(j)} \equiv \cos\p{\alpha/2}s_\x^{(j)} + \sin\p{\alpha/2}s_\y^{(j)}$.
+where $\simeq$ denotes equality up to an overall phase, and $s_\alpha^{(j)} \equiv \cos\alpha s_\x^{(j)} + \sin\alpha s_\y^{(j)}$.
 This decomposition implies
 \begin{align}
-  G_\theta = \exp\p{-i\pi S_\x^{(\theta/2)}} \exp\p{-i\pi S_\x^{(\theta)}}.
-  \label{eq:gauge_decomposition}
+  G_\theta \simeq \exp\p{i\pi S_\x^{(0)}} \exp\p{i\pi S_\x^{(\theta/2)}},
+  \label{eq:G_pi_pi}
 \end{align}
-The unitary in Eqns.~\eqref{eq:gauge_rotation} and \eqref{eq:gauge_decomposition} can therefore be implemented by a sequence of two $\pi$-pulses: one $\pi$-pulse by a clock laser with SOC angle $\phi=0$ (with Hamiltonian $\sim S_\x^{(\theta)}$ in the gauged frame), and a second $\pi$-pulse by a clock laser with SOC angle $\phi=\theta/2$ (with Hamiltonian $\sim S_\x^{(\theta/2)}$ in the gauged frame).
-This unitary can be used to reconcile the incommensurate SOC requirements of the squeezing and clock interrogation protocols.
+which can be implemented using one SOC-free ($\phi=0$) clock laser, and one clock laser with SOC angle $\phi=\theta/2$.
+Appending the two $\pi$-pulses given by Eqn.~\eqref{eq:G_pi_pi} to our squeezing protocol thus prepares a state $\ket{\tilde\xi_{\t{OAT}}^{(\theta)}}$ that is squeezed with respect to the homogenous collective spin operators $S_\x,S_\y,S_\z$ accessible to the SOC-free ($\phi=0$) clock laser used during clock interrogation.
 
-Starting with a spin-down-ploarized initial state
+As presented, the combined squeezing and clock interrogation protocols require three clock lasers in total: one without SOC ($\phi=0$), and one each for SOC angles $\phi\in\set{\theta/2,\theta}$.
+We can use $G_\theta$, however, to decompose any pulse $\exp\p{-i\v\beta\c\v S_\theta}$ into composite pulses that use only $\v S_0$ and $\v S_{\theta/2}$:
 \begin{align}
-  \ket{-\t{Z}}\equiv\p{\prod_j c_{j,\dn}^\dag}\ket{\t{vacuum}},
+  \exp\p{-i\v\beta\c\v S_\theta}
+  = G_\theta^\dag \exp\p{-i\v\beta\c\v S_0} G_\theta.
+  \label{eq:pulse_decomposition}
 \end{align}
-our OAT protocol essentially prepares the squeezed state
+The state prepared by the OAT squeezing protocol, for example, can be equivalently prepared via
 \begin{align}
-  \ket{\xi_{\t{OAT}}}
-  \equiv e^{-iH_{\t{FH}}^\theta t}
-  e^{-i\p{\pi/2} S_\x} \ket{-\t{Z}},
+  \ket{\xi_{\t{OAT}}^{(\theta)}}
+  &= e^{-i H_{\t{FH}}^{(0)}t} G_\theta^\dag
+  e^{-i\p{\pi/2}S_\x^{(0)}} G_\theta \ket{-\t{Z}} \notag \\
+  &\simeq e^{-i H_{\t{FH}}^{(0)}t} e^{-i\pi S_\x^{(\theta/2)}}
+  e^{i\p{\pi/2} S_\x^{(0)}} \ket{-\t{Z}}.
 \end{align}
-for some free evolution time $t$ and a Fermi-Hubbard Hamiltonian $H_{\t{FH}}^\theta$ exhibiting SOC with angle $\theta\ll1$, where $\theta=\phi$ in Eqn~\eqref{eq:FermiHubbard}.
-The follow-up clock interrogation protocol needs to rotate this squeezed state, most generally by some unitary $\exp\p{-i\v\alpha\c\v S}$, and then measure collective spin observables of the form (for example)
-\begin{align}
-  \bk{S_\x}_{\t{OAT}}^{\v\alpha}
-  \equiv \bk{\xi_{\t{OAT}}|e^{i\v\alpha\c\v S}
-    S_\x e^{-i\v\alpha\c\v S}|\xi_{\t{OAT}}}.
-\end{align}
-In a clock interrogation protocol with a shallow lattice, the rotation $\exp\p{-i\v\alpha\c\v S}$ and measurement of $S_\x$ must be performed with an SOC-free ($\phi=0$) clock laser, which can effectively only access the gauge-transformed operators $S_{\x,\y}^{(\theta)}$.
-We therefore expand
-\begin{align}
-  \bk{S_\x}_{\t{OAT}}^{\v\alpha}
-  &= \bk{\xi_{\t{OAT}}| G_\theta^\dag e^{i\v\alpha\c\v S^{(\theta)}}
-    S_\x^{(\theta)} e^{-i\v\alpha\c\v S^{(\theta)}}
-    G_\theta |\xi_{\t{OAT}}} \notag \\
-  &= \bk{\xi_{\t{OAT}}^\theta| e^{i\v\alpha\c\v S^{(\theta)}}
-    S_\x^{(\theta)} e^{-i\v\alpha\c\v S^{(\theta)}} |\xi_{\t{OAT}}^\theta},
-\end{align}
-where $\v S^{(\theta)}\equiv G_\theta \v S G_\theta^\dag$ and
-\begin{align}
-  \ket{\xi_{\t{OAT}}^\theta} \equiv G_\theta \ket{\xi_{\t{OAT}}}.
-\end{align}
-We are thus free to use the SOC-free clock laser during clock interrogation if we append the operation $G_\theta$ to our squeezing protocol.
-
-The last subtlety of clock interrogation and ``gauge-switching'' concerns the squeezing protocol itself.
-Experimentally implementing $G_\theta$ to enable clock interrogation requires clock lasers with SOC angles $\phi=0$ and $\phi=\theta/2$, neither of which is homogenous in our working gauge.
-Our squeezing protocols, however, need a homogenous clock laser for (i) an initial $\pi/2$ rotation of $\ket{-\t{Z}}$, (ii) spin-echo pulses throughout the squeezing protocol, and (iii) the drive used to transform OAT into TAT.
-Any clock laser pulses in the squeezing protocols can still be implemented using $G_\theta$ and the SOC-free ($\phi=0$) clock laser via the same trick used for clock interrogation, e.g.~by decomposing
-\begin{align}
-  \ket{\xi_{\t{OAT}}^\theta}
-  &= G_\theta e^{-iH_{\t{FH}}^\theta t}
-  G_\theta^\dag e^{-i\p{\pi/2} S_\x^{(\theta)}} G_\theta \ket{-\t{Z}},
-  \notag \\
-  &\simeq G_\theta e^{-iH_{\t{FH}}^\theta t}
-  G_\theta^\dag e^{-i\p{\pi/2} S_\x^{(\theta)}} \ket{-\t{Z}},
-\end{align}
-where $\simeq$ denotes equality up to a global phase, and we used the fact that the initial state $\ket{-\t{Z}}$ is an eigenstate of $G_\theta$.
-Applying a continuous drive, meanwhile, seemingly requires an extra clock laser with SOC angle $\phi=\theta$.
-In practice, it may be prohibitively difficult to properly calibrate the relative angles of three clock lasers.
+Spin-echo pulses applied throughout OAT can likewise be decomposed according to \eqref{eq:pulse_decomposition}, eliminating the need for a clock laser with SOC angle $\phi=\theta$.
+Applying a continuous drive during a squeezing protocol, however, still requires all three clock three lasers.
+If carefully tuning the relative orientations of three clock lasers proves to be too difficult in practice, converting OAT into TAT would therefore need to be done with a pulsed drive protocol, as e.g.~in Ref [\citenum{liu2011spin}].
 % note to authors: the use of a third laser is also incompatible with a 3D accordion lattice, but I (MP) decided that this point would require too much extra discussion to explain here
-In this case, converting OAT into TAT would need to be done with a pulsed drive protocol, as e.g.~in Ref.[\citenum{liu2011spin}].
 
@@
 
-In order to compute squeezing of a collective spin system, we need to compute expectation values of collective-spin operators.
+In order to compute squeezing of a collective spin system, we need to compute expectation values of collective spin operators.
 Choosing the basis $\set{\S_{\v m}}$ for these operators, where $\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$ with
 $\v m\equiv\p{m_+,m_\z,m_-}\in\mathbb{N}_0^3$, we can expand all
-collective-spin Hamiltonians in the form
+collective spin Hamiltonians in the form
 \begin{align}
@@
   \bk{\S_{\v m}}
-  &= \sum_{k\ge0} \f{t^k}{k!} \bk{\f{d^k}{dt^k} \S_{\v m}}_{t=0} \notag \\
+  &= \sum_{k\ge0} \f{t^k}{k!} \bk{\f{d^k}{dt^k} \S_{\v m}}_{t=0}
+  \notag \\
   &= \sum_{k\ge0} \f{t^k}{k!}

### 2019-05-21 soc_squeezing/collective_paper/collective_spin.tex
minor changes to collective spin paper, prior to arXiv submission

@@
   Heisenberg operators, we compute the relevant structure constants of
-  a collective operator algebra.  We demonstrate the utility of our
-  method by computing spin squeezing, two-time correlation functions,
-  and out-of-time-ordered correlators for $10^4$ spins in
+  a collective spin operator algebra.  We demonstrate the utility of
+  our method by computing spin squeezing, two-time correlation
+  functions, and out-of-time-ordered correlators for $10^4$ spins in
   strong-decoherence regimes that are otherwise inaccessible via
@@
 conservation divide the total Hilbert space of a collective spin
-system into superseletion sectors that grow only linearly with system
+system into superselection sectors that grow only linearly with system
 size $N$, thereby admitting efficient classical simulation of its
@@
 initial-time expectation values $\bk{\hat\S_{\v m}}_{t=0}$ and the
-number qof matrix elements $T_{\v m\v n}$ that we may need to
+number of matrix elements $T_{\v m\v n}$ that we may need to
 explicitly compute roughly in half.
@@
 (see Appendix \ref{sec:linear_response}).  Similar correlators have
-made appearances as order parameters that diagnose time-crystalline
+made appearances as order parameters for diagnosing time-crystalline
 phases of matter\cite{tucker2018shattered}.  Squared commutators such
@@
 \begin{acknowledgments}
-  We acknowledge helpful discussions with Robert Lewis-Swan and Kris
-  Tucker.  This work is supported by the Air Force Office of
-  Scientific Research (AFOSR) grant FA9550-18-1-0319; the AFOSR
-  Multidisciplinary University Research Initiative (MURI) grant; the
-  Defense Advanced Research Projects Agency (DARPA) and Army Research
-  Office (ARO) grant W911NF-16-1-0576 and W911NF-19-1-0210; the
-  National Science Foundation (NSF) grant PHY-1820885; JILA-NSF grant
-  PFC-173400; and the National Institute of Standards and Technology
-  (NIST).
+  We acknowledge helpful discussions with Robert Lewis-Swan, Kris
+  Tucker, Diego Barberena, and Colin Kennedy.  This work is supported
+  by the Air Force Office of Scientific Research (AFOSR) grant
+  FA9550-18-1-0319; the AFOSR Multidisciplinary University Research
+  Initiative (MURI) grant; the Defense Advanced Research Projects
+  Agency (DARPA) and Army Research Office (ARO) grant W911NF-16-1-0576
+  and W911NF-19-1-0210; the National Science Foundation (NSF) grant
+  PHY-1820885; JILA-NSF grant PFC-173400; and the National Institute
+  of Standards and Technology (NIST).
 \end{acknowledgments}
@@
 \end{align}
-The expectation values in \eqref{eq:S+_OAT}--\eqref{eq:SS+-_OAT} and
-\eqref{eq:Sz_OAT} are sufficient to compute the spin squeezing
-parameter $\xi^2$ in \eqref{eq:squeezing} at any time throughout
-evolution of the initial state $\ket\X$ under $H_{\t{OAT}}$, under
-appropriate assumptions about the relevant sources of decoherence.
+With appropriate assumptions about the relevant sources of
+decoherence, the expectation values in
+\eqref{eq:S+_OAT}--\eqref{eq:SS+-_OAT} and \eqref{eq:Sz_OAT} are
+sufficient to compute the spin squeezing parameter $\xi^2$ in
+\eqref{eq:squeezing} at any time throughout evolution of the initial
+state $\ket\X$ under $H_{\t{OAT}}$.
 
@@
 \end{align}
-where $T$ is a super-operator that generates time evolution for
+where $T$ is a superoperator that generates time evolution for
 Heisenberg operators.  In the present example, the matrix elements

### 2019-05-31 soc_squeezing/squeezing_paper/main.tex
fixed minor error in squeezing paper added two references on characterizing inter-atomic interactions added bibliography for squeezing paper

@@
 
-% The understanding and application of few-body quantum mechanics and equilibrium statistical mechanics were integral to the rapid growth in technology throughout the 20th century.
 A major frontier of contemporary physics is the understanding of non-equilibrium behaviors of many-body quantum systems, and the application of these behaviors toward the development of novel quantum technologies with untapped capabilities\cite{eisert2015quantum}.
@@
 Optical lattice clocks in particular have seen some of the most impressive developments in recent years, reaching record levels of precision ($\sim 3\times 10^{-19}$)\cite{campbell2017fermidegenerate, marti2018imaging} and accuracy ($\sim 1\times 10^{-18}$)\cite{bloom2014optical, mcgrew2018atomic}.
-These advancements required important breakthroughs, including the capability to cool and trap fermionic alkaline-earth atoms in spin-insensitive potentials\cite{takamoto2003spectroscopy, barber2006direct, ye2008quantum}; the development of ultra-coherent lasers\cite{kessler2012sub40mhzlinewidth, cole2013tenfold, matei2017mu} to fully exploit an ultra-narrow clock transition\cite{ludlow2015optical}; and, more recently, the preparation of a quantum degenerate gas in a three-dimensional (3D) optical lattice\cite{campbell2017fermidegenerate, marti2018imaging, goban2018emergence}.
+These advancements required important breakthroughs, including the capability to cool and trap fermionic alkaline-earth atoms in spin-insensitive potentials\cite{takamoto2003spectroscopy, barber2006direct, ye2008quantum}; the development of ultra-coherent lasers\cite{kessler2012sub40mhzlinewidth, cole2013tenfold, matei2017mu} to fully exploit an ultra-narrow clock transition\cite{ludlow2015optical}; the detailed characterization of inter-atomic interactions\cite{scazza2014observation, cappellini2014direct}; and, more recently, the preparation of a quantum degenerate gas in a three-dimensional (3D) optical lattice\cite{campbell2017fermidegenerate, marti2018imaging, goban2018emergence}.
 Nonetheless, all improvements in sensing capabilities to date have been based on single-particle control of internal atomic degrees of freedom.
@@
 \begin{align}
-  \hat H_{\text{FH}}^{(\phi)}/\hbar
+  \hat H_{\t{FH}}^{(\phi)}/\hbar
   &= -J \sum_j \p{e^{i\phi} \hat c_{j,\up}^\dag \hat c_{j+1,\up}

### 2019-08-28 soc_squeezing/squeezing_paper/prx/reviews/response.tex
changes to squeezing paper after PRX reviews

--- 2019-05-21 soc_squeezing/squeezing_paper/reviews/referee_response.tex
@@
 
-  We refer readers to the new appendix at the end of the caption to
+  In order to distinguish Fermi-Hubbard Hamiltonians with and without
+  SOC, we have added an explicit dependence of the Fermi-Hubbard
+  Hamiltonians in Eqns.~(1) and (2) on the SOC angle $\phi$ by
+  changing \red{$\hat H_{\t{FH}}$} and \red{$\hat H_{\t{FH,single}}$}
+  to \green{$\hat H_{\t{FH}}^{(\phi)}$} and
+  \green{$\hat H_{\t{FH,single}}^{(\phi)}$}.
+
+  We refer readers to this new appendix at the end of the caption to
   Fig.~1 on page 3, adding the text ``\green{(see Appendix D)}''; as
@@
   % TODO: make sure reference numbers are correct
+
+  We hope that these changes help clarify subtleties related to
+  interfacing our squeezing protocol with a subsequent clock
+  interrogation protocol.
 

### 2019-08-28 soc_squeezing/squeezing_paper/main.tex
changes to squeezing paper after PRX reviews

@@
-\documentclass[aps,prx,superscriptaddress,twocolumn]{revtex4-2}
+\documentclass[aps,prx,superscriptaddress,notitlepage,twocolumn,longbibliography]{revtex4-1}
 
@@
 Optical lattice clocks in particular have seen some of the most impressive developments in recent years, reaching record levels of precision ($\sim 3\times 10^{-19}$)\cite{campbell2017fermidegenerate, marti2018imaging} and accuracy ($\sim 1\times 10^{-18}$)\cite{bloom2014optical, mcgrew2018atomic}.
-These advancements required important breakthroughs, including the capability to cool and trap fermionic alkaline-earth atoms in spin-insensitive potentials\cite{takamoto2003spectroscopy, barber2006direct, ye2008quantum}; the development of ultra-coherent lasers\cite{kessler2012sub40mhzlinewidth, cole2013tenfold, matei2017mu} to fully exploit an ultra-narrow clock transition\cite{ludlow2015optical}; the detailed characterization of inter-atomic interactions\cite{scazza2014observation, cappellini2014direct}; and, more recently, the preparation of a quantum degenerate gas in a three-dimensional (3D) optical lattice\cite{campbell2017fermidegenerate, marti2018imaging, goban2018emergence}.
+These advancements required important breakthroughs, including the capability to cool and trap fermionic alkaline-earth atoms in spin-insensitive potentials\cite{takamoto2003spectroscopy, barber2006direct, ye2008quantum}; the development of ultra-coherent lasers\cite{kessler2012sub40mhzlinewidth, cole2013tenfold, matei2017mu} to fully exploit an ultra-narrow clock transition\cite{ludlow2015optical}; the detailed characterization of inter-atomic interactions\cite{scazza2014observation, cappellini2014direct, Zhang2014}; and, more recently, the preparation of a quantum degenerate gas in a three-dimensional (3D) optical lattice\cite{campbell2017fermidegenerate, marti2018imaging, goban2018emergence}.
 Nonetheless, all improvements in sensing capabilities to date have been based on single-particle control of internal atomic degrees of freedom.
@@
 However, an initially spin-polarized Fermi gas still exhibits interactions at later times due to spin-orbit coupling (SOC) that is induced by the laser that drives the clock transition (i.e.~the ``clock laser'')\cite{wall2016synthetic, kolkowitz2016spinorbitcoupled, livi2016synthetic, bromley2018dynamics}.
-Specifically, the momentum kick imparted by this laser imprints a position-dependent phase that induces inhomogenous spin precession and generates spin dephasing, thereby making atoms distinguishable and vulnerable to collisions.
+Specifically, the momentum kick imparted by this laser imprints a position-dependent phase that induces inhomogeneous spin precession and generates spin dephasing, thereby making atoms distinguishable and vulnerable to collisions.
 While a deep lattice can suppress SOC, it also intensifies the light scattering which currently limits the coherence time of the clock\cite{dorscher2018latticeinduced, goban2018emergence, hutson2019engineering}.
@@
 This scheme is made possible in the weak SOC regime by the formation of an interaction-energy gap that suppresses the SOC-induced population transfer from the exchange-symmetric Dicke manifold (spanned by spin-polarized and thus non-interacting states) to the remainder of Hilbert space.
-In this regime, the non-equilibrium dynamical processes induced by SOC result in all-to-all collective spin interactions that naturally generate spin-squeezed states in a manner robust to realistic experimental imperfections.
 To generate spin squeezing, our protocol only requires control over (i) the orientation of the clock laser and (ii) the optical lattice depth.
@@
 Additionally, we show that by applying a modulated drive from the clock laser, one can further prepare states that saturate the Heisenberg limit $\Delta\theta \sim 1/N$ for phase sensitivity\cite{kitagawa1993squeezed, ma2011quantum, degen2017quantum}.
+
+The intriguing and non-trivial observation reported here is that weakly interacting fermions in a lattice with dynamics described by the Hubbard model --- which thus obey fermionic quantum statistics and interact solely via contact collisions --- can feature all-to-all collective spin interactions that naturally generate spin-squeezed states in a manner robust to realistic experimental imperfections.
+In the spirit of quantum simulation, our work illustrates that short-range interacting systems can be used to emulate  dynamical magnetic behaviors found in arrays of long-range interacting systems (such as arrays of trapped ions\cite{Bohnet2016} or cavity QED systems\cite{Monika2010,norcia2018cavitymediated} with interactions mediated by phonons or photons, respectively), thus enhancing their potential as a metrological resource.
+Although surprising, this result is in line with the recent understandings in quantum research that explore how the interplay between well understood simple phenomena can give rise to emergent new  behaviors.
+In this direction, our work emphasizes that the precise control of interactions, which becomes possible in 3D optical lattice systems with quantized motional degrees of freedom in all three directions, is a key for the implementation of the protocols studied here, which may enable new quantum technologies.
 
@@
 
-On its own, the collective Heisenberg term ($\sim\hat{\v S}\c\hat{\v S}$) in Eqn.~\eqref{eq:SpinModel} opens an energy gap $fU$, with $f\equiv N/L$ the filling fraction of spatial modes, between the collective Dicke states $\ket{S=N/2, M_S}$ and the remainder of Hilbert space\cite{rey2008manybody, martin2013quantum, norcia2018cavitymediated} with $S<N/2$.
+On its own, the collective Heisenberg term ($\sim\hat{\v S}\c\hat{\v S}$) in Eqn.~\eqref{eq:SpinModel} opens an energy gap $fU$, with $f\equiv N/L$ the filling fraction of spatial modes, between the collective Dicke states $\ket{S=N/2, M_S}$ and the remainder of Hilbert space\cite{rey2008manybody, martin2013quantum, norcia2018cavitymediated,Smale2018} with $S<N/2$.
 Here $S$ and $M_S$ respectively label the eigenvalues of the collective spin operators $\hat{\v S}\c\hat{\v S}$ and $\hat{S}_\z$, with eigenvalues $S(S+1)$ for non-negative $S\in\set{N/2,N/2-1,\cdots}$ and $M_S\in\set{-S,-S+1,\cdots,S}$.
@@
 Here $\braket{\hat{\v S}}$ is the mean collective spin vector and $\braket{\t{var}(\hat S^\perp_\theta)}$ is the variance of spin measurements along an axis orthogonal to $\braket{\hat{\v S}}$, parameterized by the angle $\theta\in[0,2\pi)$.
+
 The above protocol concerns only the preparation of a spin-squeezed state, which would then be used as an input state for a follow-up clock interrogation protocol without SOC.
-We discuss subtleties of interfacing the squeezing protocol with a subsequent clock interrogation protocol in Appendix \ref{sec:clock_interrogation}.
+While increasing the lattice depth to turn off SOC during clock interrogation is the simplest approach, this will limit the interrogation time due to light scattering (see discussion below).
+Alternatively, it is possible to keep the same lattice depth used for the spin squeezing generation by adding a specific pulse sequence to suppress SOC.
+See details in Appendix \ref{sec:clock_interrogation}.
 
@@
 
+In realistic implementations, the Gaussian profile of the laser beams always introduces an additional effective harmonic potential that modifies the translational invariance assumed so far.
+We present a detailed discussion of the role of the harmonic trap in Appendix \ref{App:harmonic_trap}, where we demonstrate that the addition of harmonic confinement barely modifies the achievable spin squeezing with currently accessible trapping frequencies.
+We find that the existence of single-particle localized modes in the lattice with harmonic confinement\cite{Rey2005,Pupillo2006} helps to protect spin squeezing and shifts the optimal parameter window to $U/J\gtrsim 2$.
+
 The above scheme for OAT achieves optimal spin squeezing that scales as $\xi^2_{\t{opt}}\sim N^{-2/3}$ with minimal intervention, i.e.~a standard Ramsey protocol.
@@
 
-There are two general strategies for converting OAT into TAT: by use of a pulsed\cite{liu2011spin} or continuous\cite{huang2015twoaxis} drive protocol.
-For simplicity, we consider the latter in our work.
+There are two general strategies for converting OAT into TAT: by use of either a pulsed\cite{liu2011spin} or continuous\cite{huang2015twoaxis} drive protocol.
+For simplicity, we consider the latter in this  work, although the pulsed protocol could provide additional advantages, as explained at the end of Appendix \ref{sec:clock_interrogation}.
 Following the prescription in Ref.~[\citenum{huang2015twoaxis}], we use the clock laser to apply an amplitude-modulated drive $\hat H_{\t{drive}}(t)/\hbar=\Omega_0\cos(\omega t)\hat{S}_\x$.
@@
 While the OAT results in ({\bf a}) are exact, the TAT results in ({\bf b}) reflect only a lower bound on the maximum squeezing obtainable, albeit one that is likely close (within a few dB) to the actual value.
-Optimal squeezing times in the presence of decoherence are generally smaller than the corresponding times shown in Fig.~\ref{fig:optima_coherent}, as decoherence generally degrades squeezing before it reaches the decoherence-free maximum.
-Sample plots of squeezing over time for particular choices of lattice size ($\ell$) and depth ($V_0/E_{\t{R}}$) are provided in the Supplementary Material\cite{supplementary_material}.
+Optimal squeezing times in the presence of decoherence are generally smaller than the corresponding times shown in Fig.~\ref{fig:optima_coherent}, as decoherence typically degrades squeezing before it reaches the decoherence-free maximum.
+Sample plots of squeezing over time for particular choices of lattice size ($\ell$) and depth ($V_0/E_{\t{R}}$) are provided in the Supplementary Material \cite{supplementary_material}.
 }
@@
 Currently, light scattering from the lattice beams induces decoherence of the clock on a time scale of $\sim10$ seconds\cite{goban2018emergence, hutson2019engineering}, which is much shorter than the natural $^3P_0$ lifetime of $\sim160$ seconds (see Appendix \ref{sec:decoherence}).
-This limitation imposes significant constraints on achievable spin squeezing, as shown in Figure \ref{fig:optima_decoherence} where the maximal squeezing with spin decay in the OAT case was determined using exact expressions for spin correlators derived in Ref.~[\citenum{foss-feig2013nonequilibrium}], while in the TAT case these correlators were determined by solving Heisenberg equations of motion for collective spin operators (see Appendix \ref{sec:collective_simulation}).
+This limitation imposes significant constraints on achievable spin squeezing, as shown in Figure \ref{fig:optima_decoherence} where the maximal squeezing with spin decay in the OAT case was determined using exact expressions for spin correlators derived in Ref.~[\citenum{foss-feig2013nonequilibrium}], while in the TAT case these correlators were determined by solving Heisenberg equations of motion for collective spin operators\cite{perlin2019shorttime} (see Appendix \ref{sec:collective_simulation}).
 Due to the fast growth of Heisenberg operators in systems with all-to-all interactions, the latter method is not always capable of simulating up to the optimal squeezing time, and thus only provides a lower bound on the maximal squeezing theoretically obtainable via TAT.
@@
 
-The source of decoherence considered above is not fundamental, and can be avoided by, for example, using two nuclear spin levels as spin-1/2 degrees of freedom that are interrogated by far-detuned Raman transitions instead of a direct optical transition\cite{mancini2015observation}.
+In addition to light scattering, $p$-wave losses from inelastic ${}^3 P_0$ collisions\cite{martin2013quantum, Zhang2014, Bishof2011} can also degrade the maximum achievable spin squeezing, particularly in shallow lattices.
+More details on $p$-wave losses are discussed in Appendix \ref{App:inelastic_collision}, where we show that operating at lattices deeper than $8 E_{\t{R}}$ may be necessary to suppress the impact of inelastic collisions on spin squeezing, at the cost of slightly increasing  light scattering.
+
+The sources of decoherence considered above are not fundamental, and can be avoided by, for instance, using two nuclear spin levels as spin-1/2 degrees of freedom that are interrogated by far-detuned Raman transitions instead of a direct optical transition\cite{mancini2015observation}.
 The strength of SOC for Raman transitions is tunable and, moreover, the lifetimes of ground nuclear spin levels are longer than 100 seconds in the lattice\cite{goban2018emergence}.
 In this case, our protocol for preparing a squeezed state would additionally end with a coherent state transfer from nuclear to electronic degrees of freedom in order to retain metrological utility for the atomic clock.
+If, for example, the $-9/2$ and $-7/2$ nuclear spin states are used for the preparation of a squeezed state, then the collective-spin entanglement of atoms can be transferred to electronic states at the end of the spin squeezing protocol with a $\sigma^-$ polarized $\pi$ pulse.
+Such a pulse can transfer $\ket{g, I=-7/2}$ to $\ket{e, I=-9/2}$ without affecting $\ket{g, I=-9/2}$.
+
+%The only additional detail that needs to be considered for experimental implementation of our protocol is the fact that SOC needs to be suppressed during clock interrogation.
+In a shallow lattice, SOC can be turned off by use of a second laser with SOC angle $\phi=0$, and a simple pulse sequence that transforms our squeezed state into the appropriate gauge (see Appendix \ref{sec:clock_interrogation}).}
 
@@
 In the parameter regimes considered in our work, interactions alone have no effect on clock interrogation: absent of coherence between states with different net spin $S$, collective $\v S\c\v S$ interactions only generate unobservable global phases within each fixed-$S$ sector of Hilbert space.
-The remaining task to allow clock interrogation after spin squeezing is therefore to turn off SOC, which inhomogeneously detunes atomic transition frequencies by an amount $B_q$ that depends on the quasi-momentum $q$ of an atom.
+Therefore, the remaining task to allow clock interrogation after spin squeezing is to turn off SOC, which inhomogeneously detunes atomic transition frequencies by an amount $B_q$ that depends on the quasi-momentum $q$ of an atom.
 The SOC-induced axial fields $B_q\sim J\sin(\phi/2)$ depend on two tunable parameters: the tunneling rate $J$ and the SOC angle $\phi$.
@@
 Increasing the lattice depth to turn off SOC is compatible with the current clock interrogation sequence, but is incompatible with ongoing efforts to mitigate light scattering from the lattice beams (currently the primary source of decoherence in the clock; see Appendix \ref{sec:decoherence}) by use of a shallow lattice\cite{hutson2019engineering}.
-We therefore devote the rest of this section to discussing strategies for turning off SOC that are compatible with the use of shallow lattices.
+We thus devote the rest of this section to discussing strategies for turning off SOC that are compatible with using the same lattice depth for clock interrogation as the spin squeezing generation.
 
@@
 The squeezing protocol needs a clock laser with a small but nonzero SOC angle $\phi\ll1$, while the clock interrogation protocol requires a clock laser with $\phi=0$.
-Simply using one clock laser with $\phi\ne0$ for the squeezing protocol and another clock laser with $\phi=0$ for the clock interrogation protocol, however, does not resolve this discrepancy, because (i) both the squeezing and clock interrogation protocols need a clock laser that is homogeneous -- i.e.~with Hamiltonian proportional to $S_\x$, (ii) lasers with different SOC angles are homogeneous in different gauges, and (iii) spin squeezing is a gauge-dependent quantity.
-More precisely, a state that is squeezed with respect to spin operators that are homogeneous (i.e.~of the form $S_\x,S_\y,S_\z$) in a particular gauge is not necessarily squeezed with respect to spin operators that are homogeneous in a different gauge.
-To resolve the fact that our squeezing protocol prepares a state that is squeezed in the ``wrong gauge'', we will work explicitly in the ``lab gauge'' of the clock interrogation protocol, in which the Fermi-Hubbard Hamiltonian is SOC-free and the $\phi=0$ clock laser is homogenous.
-We will then construct a simple pulse sequence that transforms the inhomogeneous spin operators accessible by the $\phi\ne0$ laser into homogeneous form in our working gauge.
+Simply using one clock laser with $\phi\ne0$ for the squeezing protocol and another clock laser with $\phi=0$ for the clock interrogation protocol, however, does not resolve this discrepancy, because 
+%(i) both the squeezing and clock interrogation protocols need a clock laser that is homogeneous -- i.e.~with Hamiltonian proportional to $S_\x$,
+%(i) lasers with different SOC angles are homogeneous in different gauges, and 
+%(ii) spin squeezing is a gauge-dependent quantity.
+%More precisely,
+a state that is squeezed with respect to spin operators that are homogeneous (i.e.~of the form $S_\x,S_\y,S_\z$) in a particular gauge is not necessarily squeezed with respect to spin operators that are homogeneous in a different gauge.
+In this appendix, we will work explicitly in the ``lab gauge'' of the clock interrogation protocol, in which the Fermi-Hubbard Hamiltonian is SOC-free and the $\phi=0$ clock laser is homogeneous.
+To resolve the fact that our squeezing protocol prepares a state that is squeezed in the ``wrong gauge'', we will construct a simple pulse sequence that transforms the inhomogeneous spin operators accessible by the $\phi\ne0$ laser into a homogeneous form in our lab gauge.
 
-Starting with a spin-down-ploarized initial state
+Starting with a spin-down-polarized initial state
 \begin{align}
@@
 \end{align}
-our OAT protocol essentially prepares the state
+our OAT protocol prepares the state
 \begin{align}
@@
 \end{align}
-Defining on-site spin operators
+Defining on-site spin operators in the lab gauge 
 \begin{align}
@@
 The state $\ket{\xi_{\t{OAT}}^{(\theta)}}$ is squeezed with respect to components of the rotated collective spin vector $\v S_\theta \equiv \p{S_\x^{(\theta)},S_\y^{(\theta)},S_\z}$.
-To take advantage of the squeezing in $\ket{\xi_{\t{OAT}}^{(\theta)}}$, therefore, the clock interrogation protocol effectively needs to rotate this state by some unitary $\exp\p{-i\v\eta\c\v S_\theta}$, and then extract information about the rotation vector $\v\eta$ from collective spin observables of the form
+Therefore, to take advantage of the squeezing in $\ket{\xi_{\t{OAT}}^{(\theta)}}$, the clock interrogation protocol effectively needs to rotate this state by some unitary $\exp\p{-i\v\eta\c\v S_\theta}$, and then extract information about the rotation vector $\v\eta$ from collective spin observables of the form
 \begin{align}
@@
 where $\O_\theta$ is some product of the rotated collective spin operators in $\v S_\theta$, e.g.~$S_\x^{(\theta)}$ or $S_\x^{(\theta)} S_\y^{(\theta)}$.
-In order to turn off SOC in a shallow lattice, however, the clock interrogation protocol is restricted to performing rotations of the form $\exp\p{-i\v\eta\c\v S_0}$ and measuring homogeneous operators $\O_0$.
-We therefore seek a ``gauge-switching'' operation $G_\theta$ that maps homogenous operators $\O_0$ onto rotated operators $\O_\theta$ via $G_\theta^\dag \O_0 G_\theta = \O_\theta$.
+In order to turn off SOC during clock interrogation, however, we are restricted to performing rotations of the form $\exp\p{-i\v\eta\c\v S_0}$ and measuring homogeneous operators $\O_0$.
+We thus seek a ``gauge-switching'' operation $G_\theta$ that maps homogeneous operators $\O_0$ onto rotated operators $\O_\theta$ via $G_\theta^\dag \O_0 G_\theta = \O_\theta$.
 Equipped with $G_\theta$, we could decompose
@@
 \end{align}
-To implement $G_\theta$ with ``global'' (i.e.~non-site-selective) experimental controls, we decompose any rotation into a product of two reflections:
+To implement $G_\theta$ with ``global'' (i.e.~non-site-selective) experimental controls, we decompose each local rotation into a product of two reflections:
 \begin{align}
@@
 which can be implemented using one SOC-free ($\phi=0$) clock laser, and one clock laser with SOC angle $\phi=\theta/2$.
-Appending the two $\pi$-pulses given by Eqn.~\eqref{eq:G_pi_pi} to our squeezing protocol thus prepares a state $\ket{\tilde\xi_{\t{OAT}}^{(\theta)}}$ that is squeezed with respect to the homogenous collective spin operators $S_\x,S_\y,S_\z$ accessible to the SOC-free ($\phi=0$) clock laser used during clock interrogation.
+Appending the two $\pi$-pulses given by Eqn.~\eqref{eq:G_pi_pi} to our squeezing protocol thus prepares a state $\ket{\tilde\xi_{\t{OAT}}^{(\theta)}}$ that is squeezed with respect to the homogeneous collective spin operators $S_\x,S_\y,S_\z$ accessible to the SOC-free ($\phi=0$) clock laser used during clock interrogation.
 
@@
 Spin-echo pulses applied throughout OAT can likewise be decomposed according to \eqref{eq:pulse_decomposition}, eliminating the need for a clock laser with SOC angle $\phi=\theta$.
-Applying a continuous drive during a squeezing protocol, however, still requires all three clock three lasers.
-If carefully tuning the relative orientations of three clock lasers proves to be too difficult in practice, converting OAT into TAT would therefore need to be done with a pulsed drive protocol, as e.g.~in Ref [\citenum{liu2011spin}].
+Applying a continuous drive during a squeezing protocol, however, still requires all three clock lasers.
+If carefully tuning the relative orientations of three clock lasers proves to be too difficult in practice, converting OAT into TAT would therefore need to be done with a pulsed drive protocol, as in Ref.~[\citenum{liu2011spin}].
 % note to authors: the use of a third laser is also incompatible with a 3D accordion lattice, but I (MP) decided that this point would require too much extra discussion to explain here
@@
   \end{pmatrix},
+  \label{eq:decay_matrix}
 \end{align}
@@
 
-In order to compute squeezing of a collective spin system, we need to compute expectation values of collective spin operators.
-Choosing the basis $\set{\S_{\v m}}$ for these operators, where $\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$ with
+In order to compute squeezing of a collective spin system, we need to compute expectation values of (homogeneous) collective spin operators.
+We compute these expectation values using a method recently developed in Ref.~[\citenum{perlin2019shorttime}], and provide a short overview of the method here.
+Choosing the basis $\set{\S_{\v m}}$ for all collective spin operators, where $\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$ with
 $\v m\equiv\p{m_+,m_\z,m_-}\in\mathbb{N}_0^3$, we can expand all
@@
 \end{align}
-The evolution of a general operator $\S_{\v m}$ under a Hamiltonian of the form in Eqn.~\eqref{eq:H_general} is then given by
+The evolution of a general correlator $\bk{\S_{\v n}}$ under a Hamiltonian of the form in Eqn.~\eqref{eq:H_general} is then given by
 \begin{align}
-  \f{d}{dt} \S_{\v m}
-  &= i \sum_{\v n} h_{\v n}
-  \sp{\S_{\v n}, \S_{\v m}}_-
-  + \sum_\J \gamma_\J \mathcal D\p{\J; \S_{\v m}} \notag \\
-  &\equiv \sum_{\v n} T_{\v m\v n} \S_{\v n},
+  \f{d}{dt} \bk{\S_{\v n}}
+  &= i \sum_{\v m} h_{\v m}
+  \bk{\sp{\S_{\v m}, \S_{\v n}}_-}
+  + \sum_\J \gamma_\J \bk{\mathcal D\p{\J; \S_{\v n}}} \notag \\
+  &\equiv \sum_{\v m} \bk{\S_{\v m}} T_{\v m\v n},
 \end{align}
@@
 These matrix elements can be calculated analytically using product and commutation rules for collective spin operators.
-We can then expand Heisenberg operators in a Taylor series about $t=0$ to write
+We can then expand correlators in a Taylor series about $t=0$ to write
 \begin{align}
-  \bk{\S_{\v m}}
-  &= \sum_{k\ge0} \f{t^k}{k!} \bk{\f{d^k}{dt^k} \S_{\v m}}_{t=0}
+  \bk{\S_{\v n}}
+  &= \sum_{k\ge0} \f{t^k}{k!} \bk{\f{d^k}{dt^k} \S_{\v n}}_{t=0}
   \notag \\
   &= \sum_{k\ge0} \f{t^k}{k!}
-  \sum_{\v n} T_{\v m\v n;k} \bk{\S_{\v n}}_{t=0},
+  \sum_{\v m} T_{\v m\v n;k} \bk{\S_{\v m}}_{t=0},
   \label{eq:time_series}
@@
 where $T_{\v m\v n;k}\equiv\sp{T^k}_{\v m\v n}$ are matrix elements of the $k$-th time derivative.
-Expectation values of collective spin operators can thus be computed via the expansion in Eqn.~\eqref{eq:time_series}, which at short times can be truncated at some finite order beyond which all terms have negligible contribution to $\bk{\S_{\v m}}$.
+Expectation values of collective spin operators can thus be computed via the expansion in Eqn.~\eqref{eq:time_series}, which at short times can be truncated at some finite order beyond which all terms have negligible contribution to $\bk{\S_{\v n}}$.
+
+\section{ Effect of a harmonic confining trap}
+\label{App:harmonic_trap}
+\setcounter{figure}{0}
+
+\begin{figure*}
+\centering
+    \includegraphics[width=0.7\linewidth]{fig_harmonic.pdf}
+    \label{fig:harmonic_trap1}
+    \caption{Dynamics of non-interacting spin-orbit coupled fermions in a 1D lattice with SOC angle $\phi = \pi/50$, plus a harmonic trap with $\Omega/J = 0.01$.
+Starting with a spin-polarized cloud in $\downarrow$ ground state, an initial clock laser pulse is applied to rotate spins into $x$, and the atoms are allowed to evolve during the dark time.
+We track the dynamics of the $\uparrow$ particle density for the cases of (a) $N = 20$ and (b) $N = 60$ atoms.
+Panel (c) shows the time-averaged fluctuations of the $\uparrow$ particle density for each site index $j$ from its initial value following the Ramsey pulse; see Eqn.~\eqref{eq:time_avg_fluc}.
+For $N = 60$, we have filled all delocalized modes as well as several localized modes, resulting in a large region of no density fluctuations at the trap center.
+Panel (d) contains the eigenspectrum for a single internal state in the presence of the trap (with the index $n$ labelling the eigenvalues in order of increasing energy), where the critical mode $n_c$ dividing the spatially delocalized and localized modes is indicated  by a black dash-dotted line.
+The highest occupied mode in the $\downarrow$ ground state for $N = 20$ and $N = 60$ is indicated by the green and red solid lines, respectively.}
+    \label{fig:harmonic_trap}
+\end{figure*}
+
+\begin{figure*}
+    \includegraphics[width=0.95\linewidth]{fig_harmonic_MB.pdf}
+    \label{fig:harmonic_trap_MB1}
+    \caption{Dynamics of interacting spin-orbit  coupled fermions in a 1D lattice plus a harmonic trap for $U/J = 1$ (a), 2 (b), and 4 (c).
+For a 1D lattice with 10 sites and an SOC angle $\phi = \pi/50$, we apply a $\pi/2$ clock laser pulse to the $\downarrow$ ground state and let the system evolve during the dark time.
+In (a.i)-(c.i) we show the squeezing dynamics of the system for both $N = 10$ (solid lines) and $N = 9$ (dashed lines) for a variety of trapping strengths.
+In (a.ii)-(c.ii), we plot the time-averaged fluctuations in total particle density, $\overline{\delta n_j}$ (as in Eqn.~\eqref{eq:time_avg_fluc} but with $\hat{n}_{j,\uparrow}$ replaced by $\sum_{\alpha}\hat{n}_{j,\alpha}$).
+In (a.iii)-(c.iii), we plot the growth of the doublon population $N_d(t)$ (see Eqn.~\eqref{eq:doublon}) as a function of time, noting the absence of squeezing in the presence of a large doublon population.
+For the chosen trap strengths, the corresponding values of $n_c$ are 28 ($\Omega/J = 0.01$), 14 ($\Omega/J = 0.04$), and 6 ($\Omega/J = 0.2$).
+In panels where the results for the homogeneous case (orange curves) are not visible, they are nearly identical to the results for $\Omega/J = 0.01$ (green curves).
+Here, we utilize periodic boundary conditions to minimize finite size effects.}
+    \label{fig:harmonic_trap_MB}
+\end{figure*}
+
+Current 3D optical lattice implementations involve a harmonic confining potential, which significantly alters the underlying single-particle eigenstructure and can potentially degrade accessible squeezing within our protocol.
+In this appendix, we examine the effect of a harmonic trap on our protocol and discuss strategies to mitigate undesired effects.
+We model the trap by the addition of the term
+\begin{align}
+    \hat{H}_{\Omega} = \hbar\Omega\sum_{j,\alpha}(j-j_0)^2\hat{n}_{j,\alpha}
+\end{align}
+to our Fermi-Hubbard model (Eqn.~\eqref{eq:FermiHubbard}), where $j_0$ denotes the trap center and $\Omega = m(\omega a)^2/2\hbar$ characterizes the trap strength for atom mass $m$, trap frequency $\omega$, and lattice spacing $a$.
+In current state-of-the-art 3D $^{87}$Sr OLC implementations, values of $\omega \approx 56 \times 2\pi~\t{sec}^{-1}$ can be achieved within each 2D layer of weak SOC by utilizing in-plane lattice depths of $5 E_R$ and a lattice depth of $60 E_R$ in the axial direction, resulting in a value of $\Omega/J \approx 0.01$.
+We restrict our discussion to 1D, although for a separable 3D lattice our arguments should extend in a straightforward manner.
+
+We briefly review the structure of the single-particle eigenstates of the system, before discussing the effects on squeezing.
+In the quasi-momentum basis, the eigenfunctions $\psi_{n,\alpha}(q) = \braket{q|n,\alpha}$ are given by the $\pi$-periodic Mathieu functions, with the corresponding energies described by the Mathieu characteristic values\cite{Rey2005}.
+In the presence of SOC, using  the gauge transformation described in the main text, we obtain the relation
+\begin{gather}
+    \psi_{n,\uparrow}(q) = \psi_{n,\downarrow}(q-\phi/a).
+    \label{eq:trap_states}
+\end{gather}
+In contrast to the case of a pure harmonic potential, which generically has spatially delocalized single-particle eigenstates, the addition of a tight-binding lattice causes eigenmodes with quantum number $n$ (index $n$ labels the eigenvalues in order of increasing energy) larger than $n_c \approx 2\sqrt{2J/\Omega}$ to become localized at corresponding lattice sites.
+Therefore the  sites at a distance $n_c/2$ from the trap center with  potential energy $2\hbar J$ define  the boundary between the delocalized modes at the trap center and the high-energy localized trap edges.
+Tunneling in the region of  localized modes is typically suppressed by large potential energy differences even in the presence of SOC.
+These modes are thus largely decoupled and do not contribute to the trap center dynamics.
+On the other hand the delocalized modes may be approximated by those of a quantum harmonic oscillator with effective mass $m^{*} = \hbar/(2Ja^2)$ and frequency $\omega^* = \sqrt{4J\Omega}$.
+
+As emphasized in the main text, the key requirements for our protocol are 1) the validity of the spin model, which depends on the pinning  of particles in their initial single particle modes, and 2) the gap protection against SOC dephasing, which arises from collective spin interactions.
+Concerning the latter point, it is desirable to maintain a weak trap so as to enable a large number of delocalized modes in the trap center, which are the only type capable of contributing to the generation of squeezing.
+Though the interactions between these modes are not strictly all-to-all, they remain long-ranged, and can thus still lead to a spin-locking effect and a protective gap\cite{rey2014probing,Smale2018}.
+For $\Omega/J = 0.01$ we have $n_c = 28$, enabling $\sim10^3$ contributing modes in each 2D layer of our system.
+Concerning the validity of the spin model, from a single-particle perspective the eigenmodes  of our $\uparrow$ states will be initially displaced in quasi-momentum space from equilibrium by $\phi/a$ as per Eqn.~\eqref{eq:trap_states}, and will generally undergo dipole oscillations and not remain strictly pinned to their initial modes.
+However, as long as we ensure the displacement is small enough to guarantee  a constant density distribution across  the  trap center, the  spin model will remain valid.
+The localized modes at the trap edges can  actually help to satisfy this condition, since they can serve as a barrier against motion.
+This is demonstrated in Fig.~\ref{fig:harmonic_trap} where we show that filling all delocalized modes guarantees that the trap center maintains a constant density; we characterize this by the time-averaged fluctuations of the $\uparrow$ density at each site $j$ about its initial value following the Ramsey pulse,
+\begin{align}
+    \overline{\delta n_{j,\uparrow}} \equiv \sqrt{\lim_{t\to\infty}\frac{1}{t}\int_0^t \d\tau \bigg(\langle\hat{n}_{j,\uparrow}(\tau)\rangle - \langle\hat{n}_{j,\uparrow}(0)\rangle\bigg)^2},
+    \label{eq:time_avg_fluc}
+\end{align}
+choosing sufficiently large evolution times to ensure convergence.
+
+In the presence of interactions, an additional point of concern is that the interplay between the trap and interactions may induce resonances that enable the formation of a significant doublon population,
+\begin{align}
+    N_d(t) = \sum_j \langle\hat{n}_{j,\uparrow}(t)\hat{n}_{j,
+    \downarrow}(t) \rangle,
+    \label{eq:doublon}
+\end{align}
+which in turn may alter the density distribution and invalidate the spin model.
+Since doublon formation in the localized edges will not have consequences for our squeezing protocol, we must only ensure that doublons are not formed in the trap center, which requires $U > \Omega (n_c/2)^2 = 2J$\cite{Pupillo2006}.
+In Fig.~\ref{fig:harmonic_trap_MB}, we perform exact simulations to assess the effect of the trap on our system.
+Though restricted to small system sizes, the results demonstrate that for $U/J \lesssim 2$, the trap will always lead to a decrease of squeezing due to the formation of doublons in the trap center, while for $U/J \gtrsim 2$, we are protected from this process even for trap strengths much stronger than the experimentally relevant ones.
+
+\section{Accounting for $p$-wave inelastic collisions}
+\label{App:inelastic_collision}
+\setcounter{figure}{0}
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.4\textwidth]{rates.pdf}
+\caption{
+Both the averaged $p$-wave inelastic collision rate $\gamma$ (orange) and the ratio of this collision rate to the optimal squeezing rate $\chi_{\text{opt}}$ (blue) are suppressed as the lattice depth increases.
+%The blue line shows  the ratio of the inelastic collision rate $\gamma$ and the optimal squeezing rate $\chi_{\text{opt}}$.
+$\chi_{\text{opt}}$ is obtained by choosing SOC angles $\phi$ that saturate $\widetilde{B}/U\approx0.05$, where $\widetilde{B}$ is the variance of the SOC-induced axial field and $U$ is the two-atom on-site interaction energy.
+%The ratio decreases with increasing lattice depth.
+}
+\label{fig:inelastic_rates}
+\end{figure}
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=0.7\textwidth]{inelasticPlot.pdf}
+\caption{\textbf{Squeezing via OAT in the presence of inelastic collisions}.
+(a) For fixed particle number $N=100$, the optimal squeezing decreases as the inelastic collision rate increases.
+Panel (b) shows squeezing over time for $\gamma/\chi_{\text{opt}}=0.04$ (solid lines), which corresponds to $U/J=6$,
+and compares it with $\gamma=0$ (dashed lines) for  different particle numbers.
+Inelastic collisions prevent the growth of optimal  squeezing with  particle number. For $N=1000$, the maximum squeezing saturates to $\sim 10$dB.
+}
+\label{fig:inelastic_squeezing}
+\end{figure*}
+
+Inelastic $^{3}\text{P}_{0}$ (electronic state $e$ or $\up$) collisions are detrimental for optical lattice clocks.
+For the  nuclear-spin-polarized gas discussed in this work, $ee$ losses are only possible via the $p$-wave scattering channel since $s$-wave collisions are suppressed by Fermi statistics.
+The big advantage here compared to  prior experiments done in a 1D lattice at $\mu$K temperature\cite{martin2013quantum} is that in a Fermi  degenerate  gas loaded in a 3D optical lattice, $p$-wave losses are further suppressed by the centrifugal barrier and Pauli blocking, and only happen through a wave-function overlap between atoms at different lattice sites.
+In this appendix, we quantify the effect of $p$-wave interactions on squeezing.
+To account for $p$-wave losses, we describe the dynamics  using a master equation for the system's density matrix $\hat{\rho}$:
+\begin{align}
+\frac{d\hat{\rho}}{dt}&=-\frac{i}{\hbar}[\hat{H}_{\text{eff}},\,\hat{\rho}]+\mathcal{L}\hat{\rho}, 
+\end{align}
+where $\hat{H}_{\text{eff}}=\chi\hat{S}_z^2$ is the effective one-axis twisting Hamiltonian obtained from  the original Fermi-Hubbard Hamiltonian with spin-orbit coupling, and $\mathcal{L}$ is the Lindblad superoperator that accounts for $p$-wave $ee$ inelastic collisions.
+This latter term can be written using a pseudo-potential approximation as\cite{rey2014probing}:
+
+\begin{eqnarray}
+\mathcal{L}\hat\rho&=&\sum_{{\bm k}{\bm k'} }\Gamma_{{\bm k}{\bm k'} }
+\Big(\hat{A}_{{\bm k}{\bm k'}}\hat\rho\hat{A}_{{\bm k}{\bm k'}}^{\dagger}
+-\frac{1}{2}\big(\hat{A}_{{\bm k}{\bm k'}}^{\dagger}\hat{A}_{{\bm k}{\bm k'}}\hat\rho
++\hat\rho\hat{A}_{{\bm k}{\bm k'}}^{\dagger}\hat{A}_{{\bm k}{\bm k'}}\big)\Big),\nonumber
+\end{eqnarray} where the jump operators are $\hat{A}_{{\bm k}{\bm k'}}=\hat{c}_{{\bm k},\up}\hat{c}_{{\bm k'},\up}$, and
+$\bm k$, $\bm k'$ sum over all the populated quasi-momentum modes. The decay matrix elements $\Gamma_{{\bm k}{\bm k'}}$ are given by:
+\begin{align}
+\Gamma_{{\bm k}{\bm k'}}&=\frac{3\pi \hbar b_{\text{im}}^3 }{m}\Big(\int d{\bm r}^{\,3}
+W[\phi_{{\bm k}}({\bm r}),\phi_{{\bm k'}}({\bm r})]\Big),\end{align} where $b_{\text{im}} = 121 a_0$\cite{Zhang2014,goban2018emergence} is the $p$-wave inelastic scattering length (with $ a_0=5.29\times 10 ^{-11}$ m the Bohr radius), $\phi_{{\bm k}}({\bm r})$ is the Bloch function with quasi-momentum ${\bm k}$, and $W[\phi_{{\bm k}}({\bm r}),\phi_{{\bm k'}}({\bm r})]\equiv
+[\big({\bm \nabla}\phi^{*}_{{\bm k}}(\bm r)\big)\phi^{*}_{{\bm k'}}(\bm r)
+-\phi^{*}_{{\bm k}}(\bm r)\big({\bm \nabla}\phi^{*}_{{\bm k'}}(\bm r)\big)\big]\nonumber\\
+\cdot\big[\big({\bm \nabla}\phi_{{\bm k}}(\bm r)\big)\phi_{{\bm k'}}(\bm r)
+-\phi_{{\bm k}}(\bm r)\big({\bm \nabla}\phi_{{\bm k'}}(\bm r)\big)\big]$.
+
+In Fig.~\ref{fig:inelastic_rates} we show  the averaged decay rate $\gamma\equiv\sum_{{\bm k}{\bm k'}}\Gamma_{{\bm k}{\bm k'}}/\ell^2$, where $\ell$ is the number of lattice sites along the $x$ and $y$ axes, as a function of the  lattice depth $V_0$ along these axes.
+Here we assume  the same lattice depth  in the $z$ direction used in the main text, $V=60E_{\text{R}}$.
+As can be seen, the decay rate $\gamma$ is largely suppressed as $V_0$ increases  but is not completely negligible at shallow lattices.
+To quantify the effect of these losses on the spin squeezing generation process, we follow a similar methodology to the one described in detail in Ref.~\cite{rey2014probing}.
+The basic idea is to take advantage of the so-called Truncated-Wigner Approximation (TWA)\cite{Polkovnikok2010, Schachenmayer2015}, which allows us to capture the development of spin squeezing using semi-classical phase-space methods.
+In the  TWA the quantum dynamics are accounted for by solving mean field equations of motion supplemented by noise.
+The mean field equations are  derived by assuming that the many-body density matrix of the system can be factorized as
+$\hat{\rho}= \bigotimes_{i} \hat{\rho}({i})$, where $\hat{\rho}(i)$ is the reduced density matrix of the  particle in quasi-momentum  mode  ${\bm q}_i$ (see Eqn.~\eqref{eq:decay_matrix}).
+Under this assumption, the non-linear mean field equations are given by
+\begin{align}
+\frac{d\rho_{\up\up}(j)}{dt}&=-\sum_{j'}\Gamma_{{\bm k}_j{\bm k}_{j'}} \rho_{\up\up}(j) \rho_{\up\up}(j'), \qquad \frac{d\rho_{\dn \dn}(j)}{dt}=0\\
+\frac{d\rho_{\up\dn}(j)}{dt}&= \rho_{\up\dn}(j) \sum_{j'}\Big ( i  \chi(\rho_{\up\up}(j')-\rho_{\dn\dn}(j'))-\frac{\Gamma_{{\bm k}_j{\bm k}_{j'}}}{2} \rho_{\up\up}(j')\Big),\nonumber
+\label{eq:MeanfieldEq}
+\end{align}
+where $\rho_{\sigma\sigma'}\equiv\langle\hat{\rho}_{\sigma\sigma'}\rangle$.
+Since we are interested in the collective behavior, one can define $\rho_{\sigma\sigma'}^T=\sum_j \rho_{\sigma\sigma'}(j)$. For these observables the equations of motion simplify to
+\begin{align}
+\frac{d\rho_{\up\up}^T}{dt}&=-f\gamma  (\rho_{\up\up}^T)^2, \qquad \frac{d\rho_{\dn \dn}^T}{dt}=0\\
+\frac{d\rho_{\up\dn}^T}{dt}&= \rho_{\up\dn}^T \Big (i  \chi(\rho_{\up\up}^T-\rho_{\dn\dn}^T)-\frac{f\gamma}{2} \rho_{\up\up}^T\Big),\nonumber
+\label{eq:MeanfieldEq}
+\end{align}
+where $f\equiv N/\ell^2$ is the filling fraction.
+
+Under the TWA, one accounts for quantum fluctuations during the dynamics by  averaging over different mean field trajectories generated by sampling over different initial conditions chosen to reconstruct the Wigner function of the initial coherent spin state at $t=0$\cite{rey2014probing}.
+This method has proven to be successful in simulating quantum spin dynamics.
+Using this approach, Fig.~\ref{fig:inelastic_squeezing} shows numerical simulation results of squeezing over time in the presence of inelastic collisions.
+For shallow lattices, the effect of inelastic collisions can limit the spin squeezing to 10 dB.
+Thus, in this regime, losses are as relevant as light scattering.
+The role of inelastic interactions could be mitigated by either operating at deeper lattices as shown in Fig.
+\ref{fig:inelastic_squeezing}, or by using nuclear spin states to generate the squeezing instead of the clock states directly.
 

### 2019-08-28 soc_squeezing/squeezing_paper/prx/reviews/revision_summary.tex
changes to squeezing paper after PRX reviews

\documentclass[reprint,onecolumn,12pt]{revtex4-2}

%%% linking references
\usepackage{hyperref}
\hypersetup{
  colorlinks=true,
  linkcolor=blue,
  urlcolor=cyan,
}

%%% physics and math packages
\usepackage{physics} % general physics package
\usepackage{amssymb} % math fonts and symbols
\usepackage{bm} % for making math symbols bold
\usepackage{braket} % for nice brackets (e.g. |X>)
\usepackage{mathtools} % for \coloneqq
\usepackage{graphicx} % for figures
\graphicspath{{./figures/}} % set path for all figures

%%% shorthands to use inside math environments
\renewcommand{\t}{\text} % text in math mode
\newcommand{\f}[2]{\dfrac{#1}{#2}} % shorthand for fractions
\newcommand{\p}[1]{\left(#1\right)} % parenthesis
\renewcommand{\sp}[1]{\left[#1\right]} % square parenthesis
\renewcommand{\set}[1]{\left\{#1\right\}} % square parenthesis
\renewcommand{\c}{\cdot} % inner product
\renewcommand{\d}{\text{d}} % integration measure, e.g. d^3 x
\renewcommand{\v}{\bm} % bold symbols (e.g. for vectors)
\newcommand{\uv}[1]{\hat{\bm #1}} % unit vectors, or vector operators
\newcommand{\bk}{\Braket}
\renewcommand{\ket}{\Ket}
\renewcommand{\bra}{\Bra}

\newcommand{\B}{\mathcal{B}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\G}{\mathcal{G}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\J}{\mathcal{J}}
\renewcommand{\L}{\mathcal{L}}
\renewcommand{\O}{\mathcal{O}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\Q}{\mathcal{Q}}
\renewcommand{\S}{\mathcal{S}}

\newcommand{\z}{\text{z}}
\newcommand{\x}{\text{x}}
\newcommand{\y}{\text{y}}

\newcommand{\up}{\uparrow}
\newcommand{\dn}{\downarrow}

\usepackage{enumitem}
\setlist[enumerate]{leftmargin=*}
\usepackage[dvipsnames]{xcolor}
\newcommand{\blue}[1]{{\color{blue} #1}}
\newcommand{\red}[1]{{\color{red} #1}}
\newcommand{\green}[1]{{\color{ForestGreen} #1}}

% for numbering points in response to referees
\newcounter{point}
\newcommand{\step}{\stepcounter{point}\setcounter{enumi}{0}}

\begin{document}

Dear editors,

In this document, we summarize all revisions made to our manuscript
since its initial submission.  Revisions made in our responses to the
referee are numbered R.$X$.$Y$, corresponding to the $Y$-th revision
in point $X$ in our response.  In addition, we summarize additional
revisions made to refine wording or fix errors in the original
manuscript; these revisions are numbered M.$Z$.

Excerpts from our original manuscript are written in \red{red}, and
excerpts from the revised version of our manuscript is written in
\green{green}.  All page, equation, and reference numbers generally
refer to those in our original manuscript, unless explicitly stated
otherwise.

\section*{Revisions in response to Referee}

\begin{enumerate}[label=(R.\arabic{point}.\arabic*)]
  \step \step
\item

  \step
\item

  \step
\item

  \step
\item

  \step
\item

\end{enumerate}

\section*{Miscellaneous revisions}

\begin{enumerate}[label=(M.\arabic*)]
\item To standardize our use of language, we have changed all
  instances of the words ``\red{interrogation laser}'' and
  ``\red{external interrogation laser}'' to ``\green{clock laser}'',
  as well as all instances of ``\red{SOC phase}'' ($\phi$) with
  ``\green{SOC angle}''.  For consistency, we have also prepended
  ``\green{Eqn.}'' or ``\green{Eqns.}'' as appropriate to all equation
  references.

\item To clarify the assumptions of the spin model that become invalid
  when $U/J\lesssim2$, we have changed the text (bottom of the left
  column on page 4):

  \red{... at $U/J\gtrsim2$ mode-changing collisions start to
    become relevant and invalidate assumptions of the spin model.}

  to:

  \green{... at $U/J\gtrsim2$ mode-changing collisions start to become
    relevant and invalidate the frozen-mode assumption of the spin
    model.}

\item We fixed a typo at the top of the right column on page 4,
  replacing the text:

  \red{... our protocol should work even in this highly dopped case;
    ...}

  by:

  \green{... our protocol should work even in this highly hole-doped
    case; ...}

\item We found an error in our summaries of the results in Figure
  4(b); to fix this error, we have changed the following text (both in
  the abstract and paragraph 1 on page 7):

  \red{... $\sim12-14$ dB of spin squeezing in $\sim1$ second with
    $10^2$--$10^3$ atoms ...}

  by:

  \green{... $\sim10-14$ dB of spin squeezing in $\sim1$ second with
    $10^2$--$10^4$ atoms ...}

\item In Appendix D (page 10), we have fixed an error in identifying
  the set jump operators corresponding to single-spin decay, replacing:

  \red{... sets of jump operators $\J_-\equiv s_-^{(j)}$ and
    $\J_\z\equiv\set{s_\z^{(j)}}$...}

  by:

  \green{... sets of jump operators $\J_-\equiv\set{s_-^{(j)}}$ and
    $\J_\z\equiv\set{s_\z^{(j)}}$...}

\end{enumerate}

\end{document}

% popular summary -- beginning of second paragraph -- added word "we"

### 2019-08-30 soc_squeezing/squeezing_paper/main.tex
some changes before submission

--- 2019-05-31 soc_squeezing/squeezing_paper/main.tex
@@
-\documentclass[aps,prx,superscriptaddress,twocolumn]{revtex4-2}
+\documentclass{nature}
+
+\bibliographystyle{naturemag}
 
@@
 
-%%% text for affiliations and contributions
-\newcommand{\JILA}{JILA, National Institute of Standards and Technology and Department of Physics, University of Colorado, Boulder, CO, 80309, USA}
-\newcommand{\CTQM}{Center for Theory of Quantum Matter, University of Colorado, Boulder, CO, 80309, USA}
-\newcommand{\contrib}{\thanks{Authors P.H.~and M.A.P.~contributed equally to this work.}}
-
-\begin{document}
-
 \title{Engineering spin squeezing in a 3D optical lattice with interacting spin-orbit-coupled fermions}
 
-\author{P.~He} \contrib
-\author{M.~A.~Perlin} \contrib
-\author{S.~R.~Muleady}
-\author{R.~J.~Lewis-Swan}
-\affiliation{\JILA}
-\affiliation{\CTQM}
-\author{R.~B.~Hutson}
-\author{J.~Ye}
-\affiliation{\JILA}
-\author{A.~M.~Rey}
-\affiliation{\JILA}
-\affiliation{\CTQM}
+%%% text for affiliations and contributions
+\newcommand{\JILA}{${}^{\ref{JILA}}$}
+\newcommand{\CTQM}{${}^{\ref{CTQM}}$}
+\newcommand{\both}{${}^{\ref{JILA},\ref{CTQM}}$}
+
+\author{
+P.~He\both\footnote{Authors P.H.~and M.A.P.~contributed equally to this work.},~
+M.~A.~Perlin\both\footnotemark[\value{footnote}],~
+S.~R.~Muleady\both,~
+R.~J.~Lewis-Swan\both,~
+R.~B.~Hutson\JILA,~
+J.~Ye\JILA,~
+and A.M.~Rey\both
+}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{document}
+\maketitle
+
+\begin{affiliations}
+\item JILA, National Institute of Standards and Technology and Department of Physics, University of Colorado, Boulder, CO, 80309, USA
+\label{JILA}
+\item Center for Theory of Quantum Matter, University of Colorado, Boulder, CO, 80309, USA
+\label{CTQM}
+\end{affiliations}
 
 \begin{abstract}
-One of the most important tasks in modern quantum science is to coherently control and entangle many-body systems, and to subsequently use these systems to realize powerful quantum technologies such as quantum-enhanced sensors.
-However, many-body entangled states are difficult to prepare and preserve since internal dynamics and external noise rapidly degrade any useful entanglement.
-Here, we introduce a protocol that counterintuitively exploits inhomogeneities, a typical source of dephasing in a many-body system, in combination with interactions to generate metrologically useful and robust many-body entangled states.
-Motivated by current limitations in state-of-the-art three-dimensional (3D) optical lattice clocks (OLCs) operating at quantum degeneracy, we use local interactions in a Hubbard model with spin-orbit coupling to achieve a spin-locking effect.
-In addition to prolonging inter-particle spin coherence, spin-locking transforms the dephasing effect of spin-orbit coupling into a collective spin-squeezing process that can be further enhanced by applying a modulated drive.
-Our protocol is fully compatible with state-of-the-art 3D OLC interrogation schemes and may be used to improve their sensitivity, which is currently limited by the intrinsic quantum noise of independent atoms.
-We demonstrate that even with realistic experimental imperfections, our protocol may generate $\sim10$--$14$ dB of spin squeezing in $\sim1$ second with $\sim10^2$--$10^4$ atoms.
-This capability exemplifies a new paradigm of using driven non-equilibrium systems to overcome current limitations in quantum metrology, allowing OLCs to enter a new regime of enhanced sensing with correlated quantum states.
+%One of the most important tasks in modern quantum science is to coherently control and entangle many-body systems, and to subsequently use these systems to realize powerful quantum technologies such as quantum-enhanced sensors.
+%However, many-body entangled states are difficult to prepare and preserve since internal dynamics and external noise rapidly degrade any useful entanglement.
+%Here, we introduce a protocol that counterintuitively exploits inhomogeneities, a typical source of dephasing in a many-body system, in combination with interactions to generate metrologically useful and robust many-body entangled states.
+%Motivated by current limitations in state-of-the-art three-dimensional (3D) optical lattice clocks (OLCs) operating at quantum degeneracy, we use local interactions in a Hubbard model with spin-orbit coupling to achieve a spin-locking effect.
+%In addition to prolonging inter-particle spin coherence, spin-locking transforms the dephasing effect of spin-orbit coupling into a collective spin-squeezing process that can be further enhanced by applying a modulated drive.
+%Our protocol is fully compatible with state-of-the-art 3D OLC interrogation schemes and may be used to improve their sensitivity, which is currently limited by the intrinsic quantum noise of independent atoms.
+%We demonstrate that even with realistic experimental imperfections, our protocol may generate $\sim10$--$14$ dB of spin squeezing in $\sim1$ second with $\sim10^2$--$10^4$ atoms.
+%This capability exemplifies a new paradigm of using driven non-equilibrium systems to overcome current limitations in quantum metrology, allowing OLCs to enter a new regime of enhanced sensing with correlated quantum states.
+One of the most important tasks in modern quantum science is to coherently control and entangle many-body systems, realizing powerful quantum technologies such as quantum-enhanced sensors.
+However, many-body entangled states are difficult to prepare and preserve due to internal dynamics and external noise.
+Here, we introduce a protocol that counterintuitively exploits interactions and inhomogeneities, typical sources of dephasing, to generate metrologically useful and robust many-body entangled states.
+Our protocol is fully compatible with state-of-the-art three-dimensional (3D) optical lattice clock (OLC) interrogation schemes, and can break through limits to sensitivity set by the intrinsic quantum noise of independent atoms.
+Even with realistic experimental imperfections, our protocol may generate $\sim10$--$14$ dB of spin squeezing in $\sim1$ second with $\sim10^2$--$10^4$ atoms.
+This capability exemplifies a new paradigm of using non-equilibrium systems to overcome limitations in quantum metrology, entering a new regime of enhanced sensing with correlated quantum states.
 \end{abstract}
 
-% keywords: ultracold atoms, optical lattices, atomic clocks, metrology, spin squeezing
-
-\maketitle
-
-\section{Introduction}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Introduction}
 
@@
 Optical lattice clocks in particular have seen some of the most impressive developments in recent years, reaching record levels of precision ($\sim 3\times 10^{-19}$)\cite{campbell2017fermidegenerate, marti2018imaging} and accuracy ($\sim 1\times 10^{-18}$)\cite{bloom2014optical, mcgrew2018atomic}.
-These advancements required important breakthroughs, including the capability to cool and trap fermionic alkaline-earth atoms in spin-insensitive potentials\cite{takamoto2003spectroscopy, barber2006direct, ye2008quantum}; the development of ultra-coherent lasers\cite{kessler2012sub40mhzlinewidth, cole2013tenfold, matei2017mu} to fully exploit an ultra-narrow clock transition\cite{ludlow2015optical}; the detailed characterization of inter-atomic interactions\cite{scazza2014observation, cappellini2014direct}; and, more recently, the preparation of a quantum degenerate gas in a three-dimensional (3D) optical lattice\cite{campbell2017fermidegenerate, marti2018imaging, goban2018emergence}.
+These advancements required important breakthroughs, including the capability to cool and trap fermionic alkaline-earth atoms in spin-insensitive potentials\cite{takamoto2003spectroscopy, barber2006direct, ye2008quantum}; the development of ultra-coherent lasers\cite{kessler2012sub40mhzlinewidth, cole2013tenfold, matei2017mu} to fully exploit an ultra-narrow clock transition\cite{ludlow2015optical}; the detailed characterization of inter-atomic interactions\cite{scazza2014observation, cappellini2014direct, zhang2014spectroscopic}; and, more recently, the preparation of a quantum degenerate gas in a three-dimensional (3D) optical lattice\cite{campbell2017fermidegenerate, marti2018imaging, goban2018emergence}.
 Nonetheless, all improvements in sensing capabilities to date have been based on single-particle control of internal atomic degrees of freedom.
@@
 However, an initially spin-polarized Fermi gas still exhibits interactions at later times due to spin-orbit coupling (SOC) that is induced by the laser that drives the clock transition (i.e.~the ``clock laser'')\cite{wall2016synthetic, kolkowitz2016spinorbitcoupled, livi2016synthetic, bromley2018dynamics}.
-Specifically, the momentum kick imparted by this laser imprints a position-dependent phase that induces inhomogenous spin precession and generates spin dephasing, thereby making atoms distinguishable and vulnerable to collisions.
+Specifically, the momentum kick imparted by this laser imprints a position-dependent phase that induces inhomogeneous spin precession and generates spin dephasing, thereby making atoms distinguishable and vulnerable to collisions.
 While a deep lattice can suppress SOC, it also intensifies the light scattering which currently limits the coherence time of the clock\cite{dorscher2018latticeinduced, goban2018emergence, hutson2019engineering}.
 
-In this work, we describe a scheme that can potentially lead to metrological advances in state-of-the-art optical lattice clocks through direct use of quantum entanglement by harnessing the interplay between nominally undesirable collisions and SOC.
-This scheme is made possible in the weak SOC regime by the formation of an interaction-energy gap that suppresses the SOC-induced population transfer from the exchange-symmetric Dicke manifold (spanned by spin-polarized and thus non-interacting states) to the remainder of Hilbert space.
-In this regime, the non-equilibrium dynamical processes induced by SOC result in all-to-all collective spin interactions that naturally generate spin-squeezed states in a manner robust to realistic experimental imperfections.
+In this work, we describe a scheme that can lead to metrological advances in state-of-the-art optical lattice clocks through direct use of quantum entanglement by harnessing the interplay between nominally undesirable collisions and SOC.
+This scheme is made possible in the weak SOC regime by the formation of an interaction-energy gap that suppresses the SOC-induced population transfer from the exchange-symmetric Dicke manifold (spanned by spin-polarized, and thus non-interacting states) to the remainder of Hilbert space.
+Interactions thereby prolong inter-particle spin coherence through a spin-locking effect, which additionally transforms the dephasing effect of SOC into a collective spin squeezing process.
 To generate spin squeezing, our protocol only requires control over (i) the orientation of the clock laser and (ii) the optical lattice depth.
 These controls are straightforward to incorporate into current 3D clock interrogation sequences without sacrificing atom numbers or coherence times.
-Additionally, we show that by applying a modulated drive from the clock laser, one can further prepare states that saturate the Heisenberg limit $\Delta\theta \sim 1/N$ for phase sensitivity\cite{kitagawa1993squeezed, ma2011quantum, degen2017quantum}.
+Additionally, we show that by applying a modulated drive from the clock laser, one can further prepare states that saturate the Heisenberg limit $\Delta\theta\sim1/N$ for phase sensitivity\cite{kitagawa1993squeezed, ma2011quantum, degen2017quantum}.
 
-\section{From the Fermi-Hubbard model to one-axis and two-axis twisting}
-\label{sec:theory}
+%The intriguing and non-trivial observation reported here is that weakly interacting fermions in a lattice with dynamics described by the Hubbard model --- which obey fermionic quantum statistics and interact solely via collisions --- can feature all-to-all collective spin interactions that naturally generate spin-squeezed states in a manner robust to realistic experimental imperfections.
+%In the spirit of quantum simulation, our work illustrates that short-range interacting systems can be used to emulate  dynamical magnetic behaviors found in arrays of long-range interacting systems (such as arrays of trapped ions\cite{bohnet2016quantum} or cavity QED systems\cite{schleier-smith2010states, norcia2018cavitymediated} with interactions mediated by phonons or photons, respectively), thus enhancing their potential as a metrological resource.
+%Although surprising, this result is in line with the recent understandings in quantum research that explore how the interplay between well understood simple phenomena can give rise to emergent new behaviors.
+
+Despite an abundance of proof-of-principle experiments with entangled states\cite{degen2017quantum, pezze2018quantum}, so far only the remarkable example of LIGO\cite{aasi2013enhanced, abbott2016gw150914} has demonstrated a quantum advantage in a state-of-the-art quantum sensing or measurement system.
+%By demonstrating that the precise control of interactions, which becomes possible in the new generation of 3D optical lattice systems with fully quantized motional degrees of freedom, can naturally give rise to metrologically useful correlated many-body fermionic states, our work opens a path to not only generate entanglement but also to use it for a quantum advantage in a world-class sensor.
+The new generation of 3D optical lattice systems have fully quantized motional degrees of freedom\cite{campbell2017fermidegenerate}, allowing for precise control of collisional interactions.
+We demonstrate how these interactions can naturally give rise to metrologically useful correlated many-body fermionic states, opening a path to not only generate entanglement, but also harness it to achieve a quantum advantage in a world-class sensor.
+Such an advance will ultimately deliver gains to real-world applications including timekeeping, navigation, communication, and our understanding of the fundamental laws of nature.
+
+% Despite numerous demonstration experiments on entangled states[CITE], with the exception of LIGO\cite{aasi2013enhanced, abbott2016gw150914} no quantum advantage has yet been realized in any state-of-the-art quantum sensing or measurement system.
+% By demonstrating that the precise control of interactions, which becomes possible in the new generation of 3D optical lattice systems with quantized motional degrees of freedom in all three directions, can naturally give rise to metrologically useful correlated many-body fermionic states, our work opens a path to not only generate entanglement but also to apply it to world-class sensors which can ultimately deliver gains to real-world applications including timekeeping, navigation, communication as well as advances in our understanding of fundamental laws of nature.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Results}
+
+\subsection{Spin squeezing with the Fermi-Hubbard model.}
 
@@
 
-\begin{figure*}
+\begin{spacing}{1}
+\begin{figure}
 \centering
@@
 We then let atoms freely evolve for a variable time $t$ (with a spin-echo pulse), after which the amount of spin squeezing can be determined experimentally from global spin measurements.
-The spin-squeezed state can be used for a follow-up clock interrogation protocol (see Appendix \ref{sec:clock_interrogation}).
+The spin-squeezed state can be used for a follow-up clock interrogation protocol (see Methods).
 }
 \label{fig:protocol_schematic}
-\end{figure*}
+\end{figure}
+\end{spacing}
 
@@
   \hat H_{\t{FH}}^{(\phi)}/\hbar
-  &= -J \sum_j \p{e^{i\phi} \hat c_{j,\up}^\dag \hat c_{j+1,\up}
-  + \hat c_{j,\dn}^\dag \hat c_{j+1,\dn} + \text{h.c.}} \notag \\
-  &\qquad + U \sum_j \hat n_{j,\up} \hat n_{j,\dn},
+  = -J \sum_j \p{e^{i\phi} \hat c_{j,\up}^\dag \hat c_{j+1,\up}
+  + \hat c_{j,\dn}^\dag \hat c_{j+1,\dn} + \text{h.c.}}
+  + U \sum_j \hat n_{j,\up} \hat n_{j,\dn},
 \label{eq:FermiHubbard}
@@
 
-On its own, the collective Heisenberg term ($\sim\hat{\v S}\c\hat{\v S}$) in Eqn.~\eqref{eq:SpinModel} opens an energy gap $fU$, with $f\equiv N/L$ the filling fraction of spatial modes, between the collective Dicke states $\ket{S=N/2, M_S}$ and the remainder of Hilbert space\cite{rey2008manybody, martin2013quantum, norcia2018cavitymediated} with $S<N/2$.
+On its own, the collective Heisenberg term ($\sim\hat{\v S}\c\hat{\v S}$) in Eqn.~\eqref{eq:SpinModel} opens an energy gap $fU$, with $f\equiv N/L$ the filling fraction of spatial modes, between the collective Dicke states $\ket{S=N/2, M_S}$ and the remainder of Hilbert space\cite{rey2008manybody, martin2013quantum, norcia2018cavitymediated, smale2019observation} with $S<N/2$.
 Here $S$ and $M_S$ respectively label the eigenvalues of the collective spin operators $\hat{\v S}\c\hat{\v S}$ and $\hat{S}_\z$, with eigenvalues $S(S+1)$ for non-negative $S\in\set{N/2,N/2-1,\cdots}$ and $M_S\in\set{-S,-S+1,\cdots,S}$.
@@
 The symmetries of SOC as expressed in Eqn.~\eqref{eq:SpinModel} dictate that this treatment should yield powers of $\hat S_\z$ when projected onto the collective Dicke manifold at higher orders in perturbation theory.
-At second order in perturbation theory (see Appendix \ref{sec:derivation_OAT}), we thus find that SOC effectively yields a one-axis twisting (OAT) model widely known to generate squeezing dynamics\cite{kitagawa1993squeezed, ma2011quantum}:
+At second order in perturbation theory (see Methods), we thus find that SOC effectively yields a one-axis twisting (OAT) model widely known to generate squeezing dynamics\cite{kitagawa1993squeezed, ma2011quantum}:
 \begin{align}
@@
 Here $\braket{\hat{\v S}}$ is the mean collective spin vector and $\braket{\t{var}(\hat S^\perp_\theta)}$ is the variance of spin measurements along an axis orthogonal to $\braket{\hat{\v S}}$, parameterized by the angle $\theta\in[0,2\pi)$.
+
 The above protocol concerns only the preparation of a spin-squeezed state, which would then be used as an input state for a follow-up clock interrogation protocol without SOC.
-We discuss subtleties of interfacing the squeezing protocol with a subsequent clock interrogation protocol in Appendix \ref{sec:clock_interrogation}.
+While increasing the lattice depth to turn off SOC during clock interrogation is the simplest approach, this will limit the interrogation time due to light scattering (see discussion below).
+Alternatively, it is possible to keep the same lattice depth used for the spin squeezing generation by adding a specific pulse sequence to suppress SOC.
+See details in the Methods section.
+
+\subsection{Model validity.}
 
@@
 
-\begin{figure*}
+\begin{spacing}{1}
+\begin{figure}
 \centering
-\includegraphics[width=0.8\textwidth]{model_benchmarking.pdf}
+\includegraphics{model_benchmarking.pdf}
 \caption{{\bf Benchmarking the spin and one-axis twisting models}.
@@
 \label{fig:model_benchmarking}
-\end{figure*}
+\end{figure}
+\end{spacing}
 
@@
 
+In realistic implementations, the Gaussian profile of the laser beams always introduces an additional effective harmonic potential that modifies the translational invariance assumed so far.
+We present a detailed discussion of the role of the harmonic trap in the Supplementary Information, where we demonstrate that the addition of harmonic confinement barely modifies the achievable spin squeezing with currently accessible trapping frequencies.
+We find that the existence of single-particle localized modes in the lattice with harmonic confinement\cite{rey2005ultracold, pupillo2006extended} helps to protect spin squeezing and shifts the optimal parameter window to $U/J\gtrsim 2$.
+
+\subsection{Two-axis twisting.}
+
 The above scheme for OAT achieves optimal spin squeezing that scales as $\xi^2_{\t{opt}}\sim N^{-2/3}$ with minimal intervention, i.e.~a standard Ramsey protocol.
 Further improvements upon this scheme can be made by introducing a time-dependent driving field that transforms the OAT Hamiltonian into a two-axis twisting (TAT) one.
-While the OAT model initially generates squeezing faster than the TAT model, the squeezing generation rate of OAT (measured in dB per second) falls off with time, while the squeezing generation rate for TAT remains approximately constant until reaching Heisenberg-limited amount of spin squeezing with $\xi^2_{\t{opt}}\sim N^{-1}$\cite{kitagawa1993squeezed, supplementary_material}.
+While the OAT model initially generates squeezing faster than the TAT model, the squeezing generation rate of OAT (measured in dB per second) falls off with time, while the squeezing generation rate for TAT remains approximately constant until reaching Heisenberg-limited amount of spin squeezing with $\xi^2_{\t{opt}}\sim N^{-1}$ (see the Supplementary Information)\cite{kitagawa1993squeezed}.
 
-There are two general strategies for converting OAT into TAT: by use of a pulsed\cite{liu2011spin} or continuous\cite{huang2015twoaxis} drive protocol.
-For simplicity, we consider the latter in our work.
+There are two general strategies for converting OAT into TAT: by use of either a pulsed\cite{liu2011spin} or continuous\cite{huang2015twoaxis} drive protocol.
+For simplicity, we consider the latter in this  work, although the pulsed protocol could provide additional advantages, as explained at the end of the Methods.
 Following the prescription in Ref.~[\citenum{huang2015twoaxis}], we use the clock laser to apply an amplitude-modulated drive $\hat H_{\t{drive}}(t)/\hbar=\Omega_0\cos(\omega t)\hat{S}_\x$.
-If the modulation frequency $\omega$ satisfies $\omega\gg N\chi$ and $\J_0\p{2\Omega_0/\omega}=\pm1/3$, where $\chi$ is the OAT squeezing strength in Eqn.~\eqref{eq:H_eff} and $\J_0$ is the zero-order Bessel function of the first kind, then up to (i) an $\sim\hat{\v S}\c\hat{\v S}$ term that contributes only overall phase factors, and (ii) an $\sim\hat S_\z$ term that can be eliminated with a simple dynamical decoupling pulse sequence (see Appendix \ref{sec:dynamical_decoupling}), the effective Hamiltonian becomes $\hat H_{\t{TAT}}^{(+)}/\hbar = (\chi/3)(\hat S_\z^2-\hat S_\x^2)$ or $\hat H_{\t{TAT}}^{(-)}/\hbar = (\chi/3)(\hat S_\y^2-\hat S_\x^2)$ (see Appendix \ref{sec:derivation_TAT}), which squeezes an initial state polarized along the $y$ or $z$ axis, respectively.
+If the modulation frequency $\omega$ satisfies $\omega\gg N\chi$ and $\J_0\p{2\Omega_0/\omega}=\pm1/3$, where $\chi$ is the OAT squeezing strength in Eqn.~\eqref{eq:H_eff} and $\J_0$ is the zero-order Bessel function of the first kind, then up to (i) an $\sim\hat{\v S}\c\hat{\v S}$ term that contributes only overall phase factors, and (ii) an $\sim\hat S_\z$ term that can be eliminated with a simple dynamical decoupling pulse sequence (more details in Supplementary Information), the effective Hamiltonian becomes $\hat H_{\t{TAT}}^{(+)}/\hbar = (\chi/3)(\hat S_\z^2-\hat S_\x^2)$ or $\hat H_{\t{TAT}}^{(-)}/\hbar = (\chi/3)(\hat S_\y^2-\hat S_\x^2)$ (see Methods), which squeezes an initial state polarized along the $y$ or $z$ axis, respectively.
 
-\section{Experimental preparation of a squeezed state and practical considerations}
+\subsection{Experimental implementation and practical considerations.}
 
@@
 \label{fig:optima_coherent}
-\end{figure*}
+\end{figure}
+\end{spacing}
 
@@
 The maximally attainable amount of spin squeezing by each protocol in Fig.~\ref{fig:optima_coherent} depends only on the atom number $N$, while the shortest attainable time is determined by choosing the largest SOC angles $\phi_\x=\phi_\y\equiv\phi$ which saturate $\widetilde{B}/U\approx0.05$.
-We impose this constraint on $\widetilde{B}/U$ to ensure validity of the OAT Hamiltonian perturbatively derived in Appendix \ref{sec:derivation_OAT}\cite{supplementary_material}.
+We impose this constraint on $\widetilde{B}/U$ to ensure validity of the OAT Hamiltonian perturbatively derived in the Methods.
 
-\begin{figure*}
+\begin{spacing}{1}
+\begin{figure}
 \centering
@@
 While the OAT results in ({\bf a}) are exact, the TAT results in ({\bf b}) reflect only a lower bound on the maximum squeezing obtainable, albeit one that is likely close (within a few dB) to the actual value.
-Optimal squeezing times in the presence of decoherence are generally smaller than the corresponding times shown in Fig.~\ref{fig:optima_coherent}, as decoherence generally degrades squeezing before it reaches the decoherence-free maximum.
-Sample plots of squeezing over time for particular choices of lattice size ($\ell$) and depth ($V_0/E_{\t{R}}$) are provided in the Supplementary Material\cite{supplementary_material}.
+Optimal squeezing times in the presence of decoherence are generally smaller than the corresponding times shown in Fig.~\ref{fig:optima_coherent}, as decoherence typically degrades squeezing before it reaches the decoherence-free maximum.
+Sample plots of squeezing over time for particular choices of lattice size ($\ell$) and depth ($V_0/E_{\t{R}}$) are provided in the Supplementary Information.
 }
 \label{fig:optima_decoherence}
-\end{figure*}
+\end{figure}
+\end{spacing}
 
-Currently, light scattering from the lattice beams induces decoherence of the clock on a time scale of $\sim10$ seconds\cite{goban2018emergence, hutson2019engineering}, which is much shorter than the natural $^3P_0$ lifetime of $\sim160$ seconds (see Appendix \ref{sec:decoherence}).
-This limitation imposes significant constraints on achievable spin squeezing, as shown in Figure \ref{fig:optima_decoherence} where the maximal squeezing with spin decay in the OAT case was determined using exact expressions for spin correlators derived in Ref.~[\citenum{foss-feig2013nonequilibrium}], while in the TAT case these correlators were determined by solving Heisenberg equations of motion for collective spin operators (see Appendix \ref{sec:collective_simulation}).
+Currently, light scattering from the lattice beams induces decoherence of the clock on a time scale of $\sim10$ seconds\cite{goban2018emergence, hutson2019engineering}, which is much shorter than the natural $^3P_0$ lifetime of $\sim160$ seconds (see the Supplementary Information).
+This limitation imposes significant constraints on achievable spin squeezing, as shown in Figure \ref{fig:optima_decoherence} where the maximal squeezing with spin decay in the OAT case was determined using exact expressions for spin correlators derived in Ref.~[\citenum{foss-feig2013nonequilibrium}], while in the TAT case these correlators were determined by solving Heisenberg equations of motion for collective spin operators\cite{perlin2019shorttime} (see the Supplementary Information).
 Due to the fast growth of Heisenberg operators in systems with all-to-all interactions, the latter method is not always capable of simulating up to the optimal squeezing time, and thus only provides a lower bound on the maximal squeezing theoretically obtainable via TAT.
@@
 
-The source of decoherence considered above is not fundamental, and can be avoided by, for example, using two nuclear spin levels as spin-1/2 degrees of freedom that are interrogated by far-detuned Raman transitions instead of a direct optical transition\cite{mancini2015observation}.
+In addition to light scattering, $p$-wave losses from inelastic ${}^3 P_0$ collisions\cite{martin2013quantum, zhang2014spectroscopic, bishof2011inelastic} can also degrade the maximum achievable spin squeezing, particularly in shallow lattices.
+More details on $p$-wave losses are discussed in the Supplementary Information, where we show that operating at lattices deeper than $8 E_{\t{R}}$ may be necessary to suppress the impact of inelastic collisions on spin squeezing, at the cost of slightly increasing  light scattering.
+
+The sources of decoherence considered above are not fundamental, and can be avoided by, for instance, using two nuclear spin levels as spin-1/2 degrees of freedom that are interrogated by far-detuned Raman transitions instead of a direct optical transition\cite{mancini2015observation}.
 The strength of SOC for Raman transitions is tunable and, moreover, the lifetimes of ground nuclear spin levels are longer than 100 seconds in the lattice\cite{goban2018emergence}.
 In this case, our protocol for preparing a squeezed state would additionally end with a coherent state transfer from nuclear to electronic degrees of freedom in order to retain metrological utility for the atomic clock.
+If, for example, the $-9/2$ and $-7/2$ nuclear spin states are used for the preparation of a squeezed state, then the collective-spin entanglement of atoms can be transferred to electronic states at the end of the spin squeezing protocol with a $\sigma^-$ polarized $\pi$ pulse.
+Such a pulse can transfer $\ket{g, I=-7/2}$ to $\ket{e, I=-9/2}$ without affecting $\ket{g, I=-9/2}$.
 
-\section{Conclusions}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Discussion}
 
@@
 
-\section{Acknowledgments}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Methods}
 
-We acknowledge helpful discussions with M.~Norcia, C.~Sanner, and M.~Mamaev.
-This work is supported by the Air Force Office of Scientific Research (AFOSR) grant FA9550-18-1-0319; the AFOSR Multidisciplinary University Research Initiative (MURI) grant; the Defense Advanced Research Projects Agency (DARPA) and Army Research Office (ARO) grant W911NF-16-1-0576; the National Science Foundation (NSF) grant PHY-1820885; JILA-NSF grant PFC-173400; and the
-National Institute of Standards and Technology (NIST).
-
-\appendix
-
-% make figure numbers include the appendix letter
-\renewcommand\thefigure{\thesection\arabic{figure}}
-
-\section{Derivation of the effective one-axis-twisting model}
-\label{sec:derivation_OAT}
+\subsection{Derivation of the effective one-axis-twisting model.}
 
@@
   H = H_0 + V,
-\end{align}
-with
-\begin{align}
+  &&
   H_0 = - \f{U}{L} \v S\c\v S,
@@
 \begin{align}
-  \O V &\equiv \P_0 V \Q_0 + \Q_0 V \P_0, \\
-  \L V &\equiv \sum_{\alpha,\beta}
+  \O V \equiv \P_0 V \Q_0 + \Q_0 V \P_0,
+  &&
+  \L V \equiv \sum_{\alpha,\beta}
   \f{\op{\alpha}\O V \op{\beta}}{E_\alpha-E_\beta},
@@
   H_{\t{eff}}^{(1)} = \P_0 V \P_0,
-\end{align}
-and
-\begin{align}
+  &&
   H_{\t{eff}}^{(2)} = -\f12 \P_0 \sp{\O V,\L V}_- \P_0,
@@
 \begin{align}
-  V &= -\sum_n \p{b_n+\overline B} s_\z^{(n)} + \Omega S_\x \notag \\
-  &= -\sum_n b_n s_\z^{(n)} - \overline B S_\z + \Omega S_\x,
+  V = -\sum_n \p{b_n+\overline B} s_\z^{(n)} + \Omega S_\x
+  = -\sum_n b_n s_\z^{(n)} - \overline B S_\z + \Omega S_\x,
 \end{align}
@@
   H_{\t{eff}}^{(1)}
-  &= \P_0\p{-\sum_n b_n s_\z^{(n)} - \overline B S_\z + \Omega S_\x} \P_0 \notag \\
-  &= -\sum_n b_n \P_0 s_\z^{(n)} \P_0 - \overline B S_\z + \Omega S_\x,
+  = \P_0\p{-\sum_n b_n s_\z^{(n)} - \overline B S_\z + \Omega S_\x} \P_0
+  = -\sum_n b_n \P_0 s_\z^{(n)} \P_0 - \overline B S_\z + \Omega S_\x,
 \end{align}
@@
 Using the particle-exchange symmetry of the Dicke states, we can therefore expand
-\begin{multline}
+\begin{align}
   \sum_n b_n \P_0 s_\z^{(n)} \P_0
-  = \sum_{n,m,m'} b_n \op{m} s_\z^{(n)} \op{m'} \\
+  = \sum_{n,m,m'} b_n \op{m} s_\z^{(n)} \op{m'}
   = \sum_n b_n \sum_{m,m'} \op{m} s_\z^{(1)} \op{m'}
   = 0,
-\end{multline}
+\end{align}
 which implies
@@
 Using the fact that all spin-$z$ operators preserve the projection of total spin onto the $z$ axis, we then have that
-\begin{multline}
+\begin{align}
   H_{\t{eff}}^{(2)}
   = -\f1{fU} \sum_{m,k,n,n'} b_n b_{n'}
-  \bk{m | s_\z^{(n)} |mk} \\
-  \times \bk{mk | s_\z^{(n')} | m},
+  \bk{m | s_\z^{(n)} |mk}
+  \bk{mk | s_\z^{(n')} | m},
   \label{eq:general_H_eff_2}
-\end{multline}
+\end{align}
 where the relevant matrix elements between the Dicke states and the spin-wave states are\cite{swallows2011suppression}
@@
 which implies
-\begin{multline}
+\begin{align}
   H_{\t{eff}}^{(2)}
-  = -\f1{fU} \sum_m \f{(N/2)^2-m^2}{N^2 (N-1)} \op{m} \\
-  \times \sum_{k,n,n'} b_n b_{n'} e^{2\pi ik\p{n-n'}/N}.
-\end{multline}
+  = -\f1{fU} \sum_m \f{(N/2)^2-m^2}{N^2 (N-1)} \op{m}
+  \sum_{k,n,n'} b_n b_{n'} e^{2\pi ik\p{n-n'}/N}.
+\end{align}
 Using the fact that $\sum_nb_n=0$, we can expand
@@
   \sum_{k,n,n'} b_n b_{n'} e^{2\pi ik\p{n-n'}/N}
-  &= \sum_{n,n'} b_n b_{n'} \sum_{k=1}^{N-1} e^{2\pi ik\p{n-n'}/N} \notag \\
-  &= \sum_{n,n'} b_n b_{n'} \sum_{k=0}^{N-1} e^{2\pi ik\p{n-n'}/N},
+  = \sum_{n,n'} b_n b_{n'} \sum_{k=1}^{N-1} e^{2\pi ik\p{n-n'}/N}
+  = \sum_{n,n'} b_n b_{n'} \sum_{k=0}^{N-1} e^{2\pi ik\p{n-n'}/N},
 \end{align}
@@
   = -\f{U}{L}\v S\c\v S - \overline B S_\z + \Omega S_\x + \chi S_\z^2,
+  &&
+  \chi \equiv \f{\widetilde B^2}{\p{N-1}fU}.
   \label{eq:H_OAT}
 \end{align}
-with
-\begin{align}
-  \chi \equiv \f{\widetilde B^2}{\p{N-1}fU}.
-\end{align}
-We benchmark the validity of this effective Hamiltonian via exact simulations of the spin [Eqn.~\eqref{eq:SpinModel}] and OAT [Eqn.~\eqref{eq:H_eff}] Hamiltonians in a system of 20 spins, finding that the relative error in maximal squeezing (in dB) of the OAT model is less than 3\% when $\widetilde{B}/U<0.06$\cite{supplementary_material}.
+We benchmark the validity of this effective Hamiltonian via exact simulations of the spin [Eqn.~\eqref{eq:SpinModel}] and OAT [Eqn.~\eqref{eq:H_eff}] Hamiltonians in a system of 20 spins, finding that the relative error in maximal squeezing (in dB) of the OAT model is less than 3\% when $\widetilde{B}/U<0.06$ (see the Supplementary Information).
 
-\section{Two-axis twisting, decoherence, and the residual axial field}
-\label{sec:derivation_TAT}
+\subsection{Two-axis twisting, decoherence, and the residual axial field.}
 
@@
   e^{\pm im\beta\sin\p{\omega t}} \O
-  &= \sum_{n=-\infty}^\infty \J_n\p{\pm m\beta} e^{in\omega t} \O \notag \\
-  &\approx \J_0\p{\pm m\beta} \O
+  = \sum_{n=-\infty}^\infty \J_n\p{\pm m\beta} e^{in\omega t} \O
+  \approx \J_0\p{\pm m\beta} \O
   = \J_0\p{m\beta} \O,
@@
   H_{\t{eff}}^{(+)}
-  &= \f{\chi}{3} \p{2 S_\z^2 + S_\y^2}
-  \simeq \f{\chi}{3} \p{S_\z^2 - S_\x^2}, \\
+  = \f{\chi}{3} \p{2 S_\z^2 + S_\y^2}
+  \simeq \f{\chi}{3} \p{S_\z^2 - S_\x^2},
+  &&
   H_{\t{eff}}^{(-)}
-  &= \f{\chi}{3} \p{S_\z^2 + 2 S_\y^2}
+  = \f{\chi}{3} \p{S_\z^2 + 2 S_\y^2}
   \simeq \f{\chi}{3} \p{S_\y^2 - S_\x^2},
@@
 In a similar spirit, one can work out that single-spin operators transverse to the $x$-axis transform as
-\begin{multline}
+\begin{align}
   \tilde s_\pm
-  \equiv \f12\p{-s_\z\pm i s_\y} \\
+  \equiv \f12\p{-s_\z\pm i s_\y}
   \to U^\dag \tilde s_\pm U
@@
   \approx \J_0\p{\beta} \tilde s_\pm,
-\end{multline}
+\end{align}
 which implies that shifting into the rotating frame of the time-dependent drive takes
@@
   s_{\y,\z} \to \J_0\p{\beta} s_{\y,\z},
-  \label{eq:dec_transformation_xyz}
-\end{align}
-and
-\begin{align}
+  &&
   s_\pm
@@
   + \f12\sp{1 \mp \J_0\p{\beta}} s_-.
-  \label{eq:dec_transformation_pm}
+  \label{eq:dec_transformation_spin}
 \end{align}
-As the TAT Hamiltonians $H_{\t{eff}}^{(\pm)}$ are realized in a rotating frame, to properly account for decoherence throughout the TAT protocol one must transform jump operators according to Eqns.~\eqref{eq:dec_transformation_xyz}-\eqref{eq:dec_transformation_pm}.
+As the TAT Hamiltonians $H_{\t{eff}}^{(\pm)}$ are realized in a rotating frame, to properly account for decoherence throughout the TAT protocol one must transform jump operators according to Eqn.~\eqref{eq:dec_transformation_spin}.
 
@@
 Unlike in the case of OAT, $S_\z$ does not commute with the TAT Hamiltonians, so its effects cannot be eliminated by a single spin-echo $\pi$-pulse $\exp\p{-i\pi S_\x}$ half way through the squeezing protocol.
-Nonetheless, we find that for $N=10^2$ ($10^3$) atoms, $\sim5$ (10) $\pi$-pulses in a CPMG (Carr-Purcell-Meiboom-Gill) sequence\cite{carr1954effects, meiboom1958modified} suffice to mitigate the effects of the $S_\z$ term in the TAT protocol (see Appendix \ref{sec:dynamical_decoupling}).
+Nonetheless, we find that for $N=10^2$ ($10^3$) atoms, $\sim5$ (10) $\pi$-pulses in a CPMG (Carr-Purcell-Meiboom-Gill) sequence\cite{carr1954effects, meiboom1958modified} suffice to mitigate the effects of the $S_\z$ term in the TAT protocol (see the Supplementary Information).
 Phase control over these pulses, specifically choices of whether to apply $\exp\p{\pm i\pi S_\x}$ or $\exp\p{\pm i\pi S_\y}$ in any given $\pi$-pulse, can be used to construct XY-$n$ pulse sequences\cite{maudsley1986modified, gullion1990new} that are robust to pulse errors.
 
-\section{Dynamical decoupling in the TAT protocol}
-\label{sec:dynamical_decoupling}
-\setcounter{figure}{0}
-
-The effective Hamiltonian resulting from a perturbative treatment of SOC is (see Appendix \ref{sec:derivation_OAT})
-\begin{align}
-  H_{\t{eff}}
-  = -\f{U}{L}\v S\c\v S - \overline B S_\z + \Omega S_\x + \chi S_\z^2,
-  \label{eq:H_eff_DD}
-\end{align}
-where $U$ is a two-atom on-site interaction strength; $L$ is the number of lattice sites; $\overline B\equiv \sum_n B_n/N$ is a residual axial field determined by the occupied quai-momentum modes $\set{n}$ (with $\abs{\set{n}}=N$ atoms total); $\Omega$ is the magnitude of a driving field; and $\chi$ is an effective OAT squeezing strength.
-The effect of the $\sim\v S\c\v S$ term is to generate a relative phase between states with different total spin $S$ (where $S=N/2$ within the Dicke manifold).
-In the absence of coherent coupling between states with different total spin, therefore, the $\sim\v S\c\v S$ term has no effect on system dynamics, and we are safe to neglect it entirely.
-
-In the parameter regimes relevant to our discussions in the main text, the operator norms of $\overline{B}\hat S_\z$ and $\chi \hat S_\z^2$ in Eqn.~\eqref{eq:H_eff_DD} will typically be comparable in magnitude.
-The OAT protocol sets $\Omega=0$, and eliminates the effect of $\overline{B}\hat S_\z$ with a spin-echo $\pi$-pulse $\exp\p{-i\pi\hat S_\x}$ applied half way through the squeezing protocol.
-The TAT protocol, meanwhile, effectively takes $\chi\hat S_\z^2+\Omega \hat S_\x\to\hat H_{\t{TAT}}^{(\pm)}$ (as defined in Appendix \ref{sec:derivation_TAT}) and $\overline{B}\hat S_\z\to\J_0\p{\beta_\pm}\overline{B}\hat S_\z$, where $\J_0$ is the zero-order Bessel function of the first kind and $\beta_\pm$ is the modulation index of the amplitude-modulated driving field $\Omega$, satisfying $\J_0\p{2\beta_\pm}=\pm1/3$.
-Unlike in the case of OAT, $\hat S_\z$ does not commute with the TAT Hamiltonian, so its effect cannot be eliminated with a spin-echo.
-Nonetheless, this term can be eliminated with a dynamical decoupling pulse sequence that periodically inverts the sign of $\hat S_\z$ while preserving $\hat H_{\t{TAT}}^{(\pm)}$.
-
-\begin{figure*}
-\centering
-\includegraphics{pulsed_squeezing.pdf}
-\caption{{\bf Optimal squeezing as a function of $\pi$-pulses} applied prior to the optimal TAT squeezing time in a CPMG sequence with ({\bf a)} $N=100$ and ({\bf b}) $N=1000$ atoms.
-Results are shown for OAT, TAT, and TAT$_{\pm,\z}$, where TAT$_{\pm,\z}$ denotes squeezing via the Hamiltonian $\hat H_{\t{TAT}}^{(\pm,\z)}\equiv \hat H_{\t{TAT}}^{(\pm)}-\J_0\p{\beta_\pm}\bk{\overline B}^{\t{rms}}_f \hat S_\z$.
-Details about experimental parameters for these simulations are provided in the text.
-}
-\label{fig:pulsed_squeezing}
-\end{figure*}
-
-Fig.~\ref{fig:pulsed_squeezing} shows the maximal squeezing generated by $N=10^2$ and $10^3$ atoms via OAT, TAT, and TAT in the presence of the mean field $\J_0\p{\beta_\pm}\overline B\hat S_\z$ as a function of the number of $\pi$-pulses performed prior to the optimal TAT squeezing time.
-These pulses are applied in a CPMG sequence $\p{\tau_n/2-\pi_\x-\tau_n/2}^n$, where $\tau_n/2$ denotes Hamiltonian evolution for a time $\tau_n/2$, $\pi_\x$ denotes the application of an instantaneous $\pi$-pulse $\exp\p{-i\pi\hat S_\x}$, and $n$ is the number of pulses, such that the optimal TAT squeezing time is $t_{\t{opt}}^{\t{TAT}}=\p{\tau_n}^n$.
-The label TAT$_{\pm,\z}$ in Fig.~\ref{fig:pulsed_squeezing} denotes squeezing through the Hamiltonian $\hat H_{\t{TAT}}^{(\pm,\z)}\equiv \hat H_{\t{TAT}}^{(\pm)}-\J_0\p{\beta_\pm}\bk{\overline B}^{\t{rms}}_f\hat S_\z$, where $\bk{\overline B}^{\t{rms}}_f$ is the root-mean-square average of $\overline B$ over choices of occupied spacial modes $\set{n}$ at fixed filling $f$ of all spatial modes in the lowest Bloch band of a periodic 2D lattice.
-While the modulation index $\beta_+$ is uniquely defined by $\J_0\p{2\beta_+}=1/3$, there are two choices of $\beta_-$ for which $\J_0\p{2\beta_-}=-1/3$; we use that which minimizes $\abs{\J_0\p{\beta_-}}$.
-Fig.~\ref{fig:pulsed_squeezing} assumes an SOC angle $\phi=\pi/50$ (although results are independent of $\phi$ for $\phi\ll1$), a reduced field variance $\widetilde{B}/U=0.05$, and a filling $f=5/6$.
-Note that as the filling $f\to1$, the residual axial field vanishes ($\overline{B}\to0$), so TAT$_{\pm,\z}\to$ TAT.
-
-\section{Clock interrogation after squeezing}
-\label{sec:clock_interrogation}
+\subsection{Clock interrogation after squeezing.}
 
@@
 Here, we discuss the use of these states in a follow-up clock interrogation protocol.
-For simplicity, we restrict our discussion to the case of squeezing in 1D, as in Section \ref{sec:theory} of the main text, with the understanding that a generalization of this discussion to higher dimensions is straightforward.
+For simplicity, we restrict our discussion to the case of squeezing in 1D, as in the main text, with the understanding that a generalization of this discussion to higher dimensions is straightforward.
 
@@
 In the parameter regimes considered in our work, interactions alone have no effect on clock interrogation: absent of coherence between states with different net spin $S$, collective $\v S\c\v S$ interactions only generate unobservable global phases within each fixed-$S$ sector of Hilbert space.
-The remaining task to allow clock interrogation after spin squeezing is therefore to turn off SOC, which inhomogeneously detunes atomic transition frequencies by an amount $B_q$ that depends on the quasi-momentum $q$ of an atom.
+Therefore, the remaining task to allow clock interrogation after spin squeezing is to turn off SOC, which inhomogeneously detunes atomic transition frequencies by an amount $B_q$ that depends on the quasi-momentum $q$ of an atom.
 The SOC-induced axial fields $B_q\sim J\sin(\phi/2)$ depend on two tunable parameters: the tunneling rate $J$ and the SOC angle $\phi$.
 The simplest way to turn off SOC is thus to increase the lattice depth prior to clock interrogation, taking $J\sim B_q\to0$.
-Increasing the lattice depth to turn off SOC is compatible with the current clock interrogation sequence, but is incompatible with ongoing efforts to mitigate light scattering from the lattice beams (currently the primary source of decoherence in the clock; see Appendix \ref{sec:decoherence}) by use of a shallow lattice\cite{hutson2019engineering}.
-We therefore devote the rest of this section to discussing strategies for turning off SOC that are compatible with the use of shallow lattices.
+Increasing the lattice depth to turn off SOC is compatible with the current clock interrogation sequence, but is incompatible with ongoing efforts to mitigate light scattering from the lattice beams (currently the primary source of decoherence in the clock; see the Supplementary Information) by use of a shallow lattice\cite{hutson2019engineering}.
+We thus devote the rest of this section to discussing strategies for turning off SOC that are compatible with using the same lattice depth for clock interrogation as the spin squeezing generation.
 
@@
 \begin{align}
-  s_\z^{(j)} &\equiv \f12 \p{ c_{j,\up}^\dag c_{j,\up}
-    - c_{j,\dn}^\dag c_{j,\dn}} \\
-  s_\x^{(j)} &\equiv \f12 \p{c_{j,\up}^\dag c_{j,\dn}
-    + c_{j,\dn}^\dag c_{j,\up}} \\
-  s_\y^{(j)} &\equiv \f{i}{2} \p{c_{j,\dn}^\dag c_{j,\up}
+  s_\z^{(j)} \equiv \f12 \p{ c_{j,\up}^\dag c_{j,\up}
+    - c_{j,\dn}^\dag c_{j,\dn}},
+  &&
+  s_\x^{(j)} \equiv \f12 \p{c_{j,\up}^\dag c_{j,\dn}
+    + c_{j,\dn}^\dag c_{j,\up}},
+  &&
+  s_\y^{(j)} \equiv \f{i}{2} \p{c_{j,\dn}^\dag c_{j,\up}
     - c_{j,\up}^\dag c_{j,\dn}},
@@
 \begin{align}
-  S_\x^{(\theta)} &\equiv \sum_j \p{\cos\p{\theta j} s_\x^{(j)}
+  S_\x^{(\theta)} \equiv \sum_j \p{\cos\p{\theta j} s_\x^{(j)}
   + \sin\p{\theta j} s_\y^{(j)}},
-  \label{eq:S_x_theta} \\
-  S_\y^{(\theta)} &\equiv \sum_j \p{\cos\p{\theta j} s_\y^{(j)}
+  &&
+  S_\y^{(\theta)} \equiv \sum_j \p{\cos\p{\theta j} s_\y^{(j)}
   - \sin\p{\theta j} s_\x^{(j)}}.
-  \label{eq:S_y_theta}
+  \label{eq:S_xy_theta}
 \end{align}
 The state $\ket{\xi_{\t{OAT}}^{(\theta)}}$ is squeezed with respect to components of the rotated collective spin vector $\v S_\theta \equiv \p{S_\x^{(\theta)},S_\y^{(\theta)},S_\z}$.
-To take advantage of the squeezing in $\ket{\xi_{\t{OAT}}^{(\theta)}}$, therefore, the clock interrogation protocol effectively needs to rotate this state by some unitary $\exp\p{-i\v\eta\c\v S_\theta}$, and then extract information about the rotation vector $\v\eta$ from collective spin observables of the form
+Therefore, to take advantage of the squeezing in $\ket{\xi_{\t{OAT}}^{(\theta)}}$, the clock interrogation protocol effectively needs to rotate this state by some unitary $\exp\p{-i\v\eta\c\v S_\theta}$, and then extract information about the rotation vector $\v\eta$ from collective spin observables of the form
 \begin{align}
@@
   \bk{\O_\theta}_{\t{OAT}}^{\v\eta}
-  &= \bk{\xi_{\t{OAT}}^{(\theta)}| G_\theta^\dag e^{i\v\eta\c\v S_0}
+  = \bk{\xi_{\t{OAT}}^{(\theta)}| G_\theta^\dag e^{i\v\eta\c\v S_0}
     \O_0 e^{-i\v\eta\c\v S_0} G_\theta |\xi_{\t{OAT}}^{(\theta)}}
-  \notag \\
-  &= \bk{\tilde \xi_{\t{OAT}}^{(\theta)}| e^{i\v\eta\c\v S_0}
+  = \bk{\tilde \xi_{\t{OAT}}^{(\theta)}| e^{i\v\eta\c\v S_0}
     \O_0 e^{-i\v\eta\c\v S_0} |\tilde \xi_{\t{OAT}}^{(\theta)}}
@@
 
-Given the definitions of the rotated spin operators $S_\x^{(\theta)},S_\x^{(\theta)}$ in Eqns.~\eqref{eq:S_x_theta}, \eqref{eq:S_y_theta}, a suitable candidate for the gauge-switching operator $G_\theta$ is the site-dependent rotation
+Given the definitions of the rotated spin operators $S_\x^{(\theta)},S_\x^{(\theta)}$ in Eqn.~\eqref{eq:S_xy_theta}, a suitable candidate for the gauge-switching operator $G_\theta$ is the site-dependent rotation
 \begin{align}
@@
   \ket{\xi_{\t{OAT}}^{(\theta)}}
-  &= e^{-i H_{\t{FH}}^{(0)}t} G_\theta^\dag
-  e^{-i\p{\pi/2}S_\x^{(0)}} G_\theta \ket{-\t{Z}} \notag \\
-  &\simeq e^{-i H_{\t{FH}}^{(0)}t} e^{-i\pi S_\x^{(\theta/2)}}
+  = e^{-i H_{\t{FH}}^{(0)}t} G_\theta^\dag
+  e^{-i\p{\pi/2}S_\x^{(0)}} G_\theta \ket{-\t{Z}}
+  \simeq e^{-i H_{\t{FH}}^{(0)}t} e^{-i\pi S_\x^{(\theta/2)}}
   e^{i\p{\pi/2} S_\x^{(0)}} \ket{-\t{Z}}.
@@
 Spin-echo pulses applied throughout OAT can likewise be decomposed according to \eqref{eq:pulse_decomposition}, eliminating the need for a clock laser with SOC angle $\phi=\theta$.
-Applying a continuous drive during a squeezing protocol, however, still requires all three clock three lasers.
-If carefully tuning the relative orientations of three clock lasers proves to be too difficult in practice, converting OAT into TAT would therefore need to be done with a pulsed drive protocol, as e.g.~in Ref [\citenum{liu2011spin}].
-% note to authors: the use of a third laser is also incompatible with a 3D accordion lattice, but I (MP) decided that this point would require too much extra discussion to explain here
+Applying a continuous drive during a squeezing protocol, however, still requires all three clock lasers.
+If carefully tuning the relative orientations of three clock lasers proves to be too difficult in practice, converting OAT into TAT would therefore need to be done with a pulsed drive protocol, as in Ref.~[\citenum{liu2011spin}].
 
-\section{Decoherence in the 3D $^{87}$Sr optical lattice clock}
-\label{sec:decoherence}
+\subsection{Data availability.}
 
-Currently, light scattering from lattice beams in the 3D $^{87}$Sr optical lattice clock induces decoherence on a time scale of $\sim$10 seconds\cite{goban2018emergence, hutson2019engineering}.
-This decoherence acts identically on all atoms in an uncorrelated manner, and can be understood by considering the density operator $\rho$ for a single atom, with effective spin states $\dn$ and $\up$ respectively corresponding to the ${}^1\t{S}_0$ and ${}^3\t{P}_0$ electronic states.
-Empirically, the effect of decoherence after a time $t$ within the $\set{\dn,\up}$ subspace of a single atom is to take $\rho\to\rho\p{t}$ with $\rho\p{0}\equiv\rho$ and
-\begin{align}
-  \rho\p{t} \coloneqq
-  \begin{pmatrix}
-  \rho_{\up\up} e^{-\Gamma_{\up\up}t} &&
-  \rho_{\up\dn} e^{-\Gamma_{\up\dn}t} \\
-  \rho_{\up\dn}^* e^{-\Gamma_{\up\dn}t} &&
-  \rho_{\dn\dn} + \p{1-e^{-\Gamma_{\up\up}t}} \rho_{\up\up}
-  \end{pmatrix},
-\end{align}
-where $\Gamma_{\up\up}\approx\Gamma_{\up\dn}\approx\Gamma=0.1~\t{sec}^{-1}$ are respectively decay rates for $\rho_{\up\up}$ and $\rho_{\up\dn}$.
-This form of decoherence can be effectively modeled by decay and dephasing of individual spins (respectively denoted $\Gamma_{\t{ud}}$ and $\Gamma_{\t{el}}$ in Ref.~[\citenum{foss-feig2013nonequilibrium}]) at rates $\Gamma$.
-In the language of the section that follows, we would say that this decoherence is captured by the sets of jump operators $\J_-\equiv\set{s_-^{(j)}}$ and $\J_\z\equiv\set{s_\z^{(j)}}$ with corresponding decoherence rates $\gamma_-=\gamma_\z=\Gamma$.
+All data generated and analysed for this study are available from the corresponding authors on reasonable request.
 
-\section{Solving Heisenberg equations of motion for collective spin systems}
-\label{sec:collective_simulation}
-
-In order to compute squeezing of a collective spin system, we need to compute expectation values of collective spin operators.
-Choosing the basis $\set{\S_{\v m}}$ for these operators, where $\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$ with
-$\v m\equiv\p{m_+,m_\z,m_-}\in\mathbb{N}_0^3$, we can expand all
-collective spin Hamiltonians in the form
-\begin{align}
-  H = \sum_{\v m} h_{\v m} \S_{\v m}.
-  \label{eq:H_general}
-\end{align}
-The evolution of a general operator $\S_{\v m}$ under a Hamiltonian of the form in Eqn.~\eqref{eq:H_general} is then given by
-\begin{align}
-  \f{d}{dt} \S_{\v m}
-  &= i \sum_{\v n} h_{\v n}
-  \sp{\S_{\v n}, \S_{\v m}}_-
-  + \sum_\J \gamma_\J \mathcal D\p{\J; \S_{\v m}} \notag \\
-  &\equiv \sum_{\v n} T_{\v m\v n} \S_{\v n},
-\end{align}
-where $\sp{X,Y}_\pm\equiv XY\pm YX$; $\J$ is a set of jump operators with corresponding decoherence rate $\gamma_\J$; the decoherence operator $\mathcal D$ is defined by
-\begin{align}
-  \mathcal{D}\p{\J;\O}
-  \equiv \sum_{J\in\J}\p{J^\dag \O J - \f12\sp{J^\dag J,\O}_+};
-\end{align}
-and $T_{\v m\v n}$ is a matrix element of the time derivative operator $T\equiv d/dt$.
-These matrix elements can be calculated analytically using product and commutation rules for collective spin operators.
-We can then expand Heisenberg operators in a Taylor series about $t=0$ to write
-\begin{align}
-  \bk{\S_{\v m}}
-  &= \sum_{k\ge0} \f{t^k}{k!} \bk{\f{d^k}{dt^k} \S_{\v m}}_{t=0}
-  \notag \\
-  &= \sum_{k\ge0} \f{t^k}{k!}
-  \sum_{\v n} T_{\v m\v n;k} \bk{\S_{\v n}}_{t=0},
-  \label{eq:time_series}
-\end{align}
-where $T_{\v m\v n;k}\equiv\sp{T^k}_{\v m\v n}$ are matrix elements of the $k$-th time derivative.
-Expectation values of collective spin operators can thus be computed via the expansion in Eqn.~\eqref{eq:time_series}, which at short times can be truncated at some finite order beyond which all terms have negligible contribution to $\bk{\S_{\v m}}$.
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\bibliographystyle{naturemag}
 \bibliography{main.bib}
 
+\subsection{Acknowledgements.}
+
+We acknowledge helpful discussions with M.~Norcia, C.~Sanner, and M.~Mamaev.
+This work is supported by the Air Force Office of Scientific Research (AFOSR) grant FA9550-18-1-0319; the AFOSR Multidisciplinary University Research Initiative (MURI) grant; the Defense Advanced Research Projects Agency (DARPA) and Army Research Office (ARO) grant W911NF-16-1-0576; the National Science Foundation (NSF) grant PHY-1820885; JILA-NSF grant PFC-173400; and the
+National Institute of Standards and Technology (NIST).
+
+\subsection{Competing interests.}
+
+The authors declare no competing financial interests.
+
+\subsection{Correspondence.}
+
+Correspondence and requests for materials should be addressed to M.A.~Perlin~(email: mika.perlin@gmail.com) and A.M.~Rey~(email: arey@jilau1.colorado.edu).
+
 \end{document}

### 2019-08-30 soc_squeezing/squeezing_paper/supplementary.tex
some changes before submission

@@
-\documentclass[prx,superscriptaddress,12pt]{revtex4-2}
+\documentclass{nature}
+
+\bibliographystyle{naturemag}
 
@@
 
-%%% text for affiliations and contributions
-\newcommand{\JILA}{JILA, National Institute of Standards and Technology and Department of Physics, University of Colorado, Boulder, CO, 80309, USA}
-\newcommand{\CTQM}{Center for Theory of Quantum Matter, University of Colorado, Boulder, CO, 80309, USA}
-\newcommand{\contrib}{\thanks{Authors P.H.~and M.A.P.~contributed equally to this work.}}
-
 %%% add "S" prefix to equation and figure numbers
@@
 
-%%% for referencing files in the main text
-\usepackage{xr}
-\externaldocument[M-]{main}
+\title{Supplementary Information for ``Engineering spin squeezing in a 3D optical lattice with interacting spin-orbit-coupled fermions''}
+
+%%% text for affiliations and contributions
+\newcommand{\JILA}{${}^{\ref{JILA}}$}
+\newcommand{\CTQM}{${}^{\ref{CTQM}}$}
+\newcommand{\both}{${}^{\ref{JILA},\ref{CTQM}}$}
+
+\author{
+P.~He\both\footnote{Authors P.H.~and M.A.P.~contributed equally to this work.},~
+M.~A.~Perlin\both\footnotemark[\value{footnote}],~
+S.~R.~Muleady\both,~
+R.~J.~Lewis-Swan\both,~
+R.~B.~Hutson\JILA,~
+J.~Ye\JILA,~
+and A.M.~Rey\both
+}
 
 \begin{document}
-
-\title{Supplementary Material for ``Engineering spin squeezing in a 3D optical lattice with interacting spin-orbit-coupled fermions''}
-
-\author{P.~He} \contrib
-\author{M.A.~Perlin} \contrib
-\author{S.R.~Muleady}
-\author{R.J.~Lewis-Swan}
-\affiliation{\JILA}
-\affiliation{\CTQM}
-\author{R.B.~Hutson}
-\author{J.~Ye}
-\affiliation{\JILA}
-\author{A.M.~Rey}
-\affiliation{\JILA}
-\affiliation{\CTQM}
-
 \maketitle
 
-\section{Benchmarking}
+\begin{affiliations}
+\item JILA, National Institute of Standards and Technology and Department of Physics, University of Colorado, Boulder, CO, 80309, USA
+\label{JILA}
+\item Center for Theory of Quantum Matter, University of Colorado, Boulder, CO, 80309, USA
+\label{CTQM}
+\end{affiliations}
+
+\section*{Benchmarking the one-axis twisting model}
 
@@
 
-In principle, spin-changing decoherence compromises the validity of the OAT model, as its perturbative derivation in Appendix \ref{M-sec:derivation_OAT} of the main text relies on spin population remaining primarily within the Dicke manifold.
+In principle, spin-changing decoherence compromises the validity of the OAT model, as its perturbative derivation in the Methods of the main text relies on spin population remaining primarily within the Dicke manifold.
 This assumption breaks down in the presence of, for example, spontaneous emission, which transfers population outside of the Dicke manifold.
@@
 \label{fig:relative_error}
-\end{figure*}
+\end{figure}
+\end{spacing}
 
-\begin{figure*}
+\begin{spacing}{1}
+\begin{figure}
 \centering
@@
 \label{fig:benchmarking_decay}
-\end{figure*}
+\end{figure}
+\end{spacing}
 
-\section{Time-series of squeezing via OAT and TAT}
+\section*{Effect of a harmonic confining trap}
+
+\begin{spacing}{1}
+\begin{figure}
+\centering
+    \includegraphics[width=0.7\linewidth]{fig_harmonic.pdf}
+    \label{fig:harmonic_trap1}
+    \caption{Dynamics of non-interacting spin-orbit coupled fermions in a 1D lattice with SOC angle $\phi = \pi/50$, plus a harmonic trap with $\Omega/J = 0.01$.
+Starting with a spin-polarized cloud in $\downarrow$ ground state, an initial clock laser pulse is applied to rotate spins into $x$, and the atoms are allowed to evolve during the dark time.
+We track the dynamics of the $\uparrow$ particle density for the cases of (a) $N = 20$ and (b) $N = 60$ atoms.
+Panel (c) shows the time-averaged fluctuations of the $\uparrow$ particle density for each site index $j$ from its initial value following the Ramsey pulse; see Eqn.~\eqref{eq:time_avg_fluc}.
+For $N = 60$, we have filled all delocalized modes as well as several localized modes, resulting in a large region of no density fluctuations at the trap center.
+Panel (d) contains the eigenspectrum for a single internal state in the presence of the trap (with the index $n$ labelling the eigenvalues in order of increasing energy), where the critical mode $n_c$ dividing the spatially delocalized and localized modes is indicated  by a black dash-dotted line.
+The highest occupied mode in the $\downarrow$ ground state for $N = 20$ and $N = 60$ is indicated by the green and red solid lines, respectively.}
+    \label{fig:harmonic_trap}
+\end{figure}
+\end{spacing}
+
+\begin{spacing}{1}
+\begin{figure}
+    \includegraphics[width=0.95\linewidth]{fig_harmonic_MB.pdf}
+    \label{fig:harmonic_trap_MB1}
+    \caption{Dynamics of interacting spin-orbit  coupled fermions in a 1D lattice plus a harmonic trap for $U/J = 1$ (a), 2 (b), and 4 (c).
+For a 1D lattice with 10 sites and an SOC angle $\phi = \pi/50$, we apply a $\pi/2$ clock laser pulse to the $\downarrow$ ground state and let the system evolve during the dark time.
+In (a.i)-(c.i) we show the squeezing dynamics of the system for both $N = 10$ (solid lines) and $N = 9$ (dashed lines) for a variety of trapping strengths.
+In (a.ii)-(c.ii), we plot the time-averaged fluctuations in total particle density, $\overline{\delta n_j}$ (as in Eqn.~\eqref{eq:time_avg_fluc} but with $\hat{n}_{j,\uparrow}$ replaced by $\sum_{\alpha}\hat{n}_{j,\alpha}$).
+In (a.iii)-(c.iii), we plot the growth of the doublon population $N_d(t)$ (see Eqn.~\eqref{eq:doublon}) as a function of time, noting the absence of squeezing in the presence of a large doublon population.
+For the chosen trap strengths, the corresponding values of $n_c$ are 28 ($\Omega/J = 0.01$), 14 ($\Omega/J = 0.04$), and 6 ($\Omega/J = 0.2$).
+In panels where the results for the homogeneous case (orange curves) are not visible, they are nearly identical to the results for $\Omega/J = 0.01$ (green curves).
+Here, we utilize periodic boundary conditions to minimize finite size effects.}
+    \label{fig:harmonic_trap_MB}
+\end{figure}
+\end{spacing}
+
+Current 3D optical lattice implementations involve a harmonic confining potential, which significantly alters the underlying single-particle eigenstructure and can potentially degrade accessible squeezing within our protocol.
+In this appendix, we examine the effect of a harmonic trap on our protocol and discuss strategies to mitigate undesired effects.
+We model the trap by the addition of the term
+\begin{align}
+  \hat{H}_{\Omega}
+  = \Omega\sum_{j,\alpha}(j-j_0)^2\hat{n}_{j,\alpha}
+\end{align}
+to our Fermi-Hubbard model, where $j_0$ denotes the trap center and $\Omega = m(\omega a)^2/2$ characterizes the trap strength for atom mass $m$, trap frequency $\omega$, and lattice spacing $a$.
+In current state-of-the-art 3D $^{87}$Sr OLC implementations, values of $\omega \approx 56 \times 2\pi~\t{sec}^{-1}$ can be achieved within each 2D layer of weak SOC by utilizing in-plane lattice depths of $5 E_R$ and a lattice depth of $60 E_R$ in the axial direction, resulting in a value of $\Omega/J \approx 0.01$.
+We restrict our discussion to 1D, although for a separable 3D lattice our arguments should extend in a straightforward manner.
+
+We briefly review the structure of the single-particle eigenstates of the system, before discussing the effects on squeezing.
+In the quasi-momentum basis, the eigenfunctions $\psi_{n,\alpha}(q) = \braket{q|n,\alpha}$ are given by the $\pi$-periodic Mathieu functions, with the corresponding energies described by the Mathieu characteristic values\cite{rey2005ultracold}.
+In the presence of SOC, using  the gauge transformation described in the main text, we obtain the relation
+\begin{gather}
+    \psi_{n,\uparrow}(q) = \psi_{n,\downarrow}(q-\phi/a).
+    \label{eq:trap_states}
+\end{gather}
+In contrast to the case of a pure harmonic potential, which generically has spatially delocalized single-particle eigenstates, the addition of a tight-binding lattice causes eigenmodes with quantum number $n$ (index $n$ labels the eigenvalues in order of increasing energy) larger than $n_c \approx 2\sqrt{2J/\Omega}$ to become localized at corresponding lattice sites.
+Therefore the  sites at a distance $n_c/2$ from the trap center with  potential energy $2J$ define  the boundary between the delocalized modes at the trap center and the high-energy localized trap edges.
+Tunneling in the region of  localized modes is typically suppressed by large potential energy differences even in the presence of SOC.
+These modes are thus largely decoupled and do not contribute to the trap center dynamics.
+On the other hand the delocalized modes may be approximated by those of a quantum harmonic oscillator with effective mass $m^{*}=1/(2Ja^2)$ and frequency $\omega^* = \sqrt{4J\Omega}$.
+
+As emphasized in the main text, the key requirements for our protocol are 1) the validity of the spin model, which depends on the pinning  of particles in their initial single particle modes, and 2) the gap protection against SOC dephasing, which arises from collective spin interactions.
+Concerning the latter point, it is desirable to maintain a weak trap so as to enable a large number of delocalized modes in the trap center, which are the only type capable of contributing to the generation of squeezing.
+Though the interactions between these modes are not strictly all-to-all, they remain long-ranged, and can thus still lead to a spin-locking effect and a protective gap\cite{rey2014probing, smale2019observation}.
+For $\Omega/J = 0.01$ we have $n_c = 28$, enabling $\sim10^3$ contributing modes in each 2D layer of our system.
+Concerning the validity of the spin model, from a single-particle perspective the eigenmodes  of our $\uparrow$ states will be initially displaced in quasi-momentum space from equilibrium by $\phi/a$ as per Eqn.~\eqref{eq:trap_states}, and will generally undergo dipole oscillations and not remain strictly pinned to their initial modes.
+However, as long as we ensure the displacement is small enough to guarantee  a constant density distribution across  the  trap center, the  spin model will remain valid.
+The localized modes at the trap edges can  actually help to satisfy this condition, since they can serve as a barrier against motion.
+This is demonstrated in Fig.~\ref{fig:harmonic_trap} where we show that filling all delocalized modes guarantees that the trap center maintains a constant density; we characterize this by the time-averaged fluctuations of the $\uparrow$ density at each site $j$ about its initial value following the Ramsey pulse,
+\begin{align}
+  \overline{\delta n_{j,\uparrow}}
+  \equiv \sqrt{\lim_{t\to\infty}\frac{1}{t}
+    \int_0^t \d\tau \bigg(\langle\hat{n}_{j,\uparrow}(\tau)\rangle
+    - \langle\hat{n}_{j,\uparrow}(0)\rangle\bigg)^2},
+    \label{eq:time_avg_fluc}
+\end{align}
+choosing sufficiently large evolution times to ensure convergence.
+
+In the presence of interactions, an additional point of concern is that the interplay between the trap and interactions may induce resonances that enable the formation of a significant doublon population,
+\begin{align}
+    N_d(t) = \sum_j \langle\hat{n}_{j,\uparrow}(t)\hat{n}_{j,
+    \downarrow}(t) \rangle,
+    \label{eq:doublon}
+\end{align}
+which in turn may alter the density distribution and invalidate the spin model.
+Since doublon formation in the localized edges will not have consequences for our squeezing protocol, we must only ensure that doublons are not formed in the trap center, which requires $U > \Omega (n_c/2)^2 = 2J$\cite{pupillo2006extended}.
+In Fig.~\ref{fig:harmonic_trap_MB}, we perform exact simulations to assess the effect of the trap on our system.
+Though restricted to small system sizes, the results demonstrate that for $U/J \lesssim 2$, the trap will always lead to a decrease of squeezing due to the formation of doublons in the trap center, while for $U/J \gtrsim 2$, we are protected from this process even for trap strengths much stronger than the experimentally relevant ones.
+
+\section*{Dynamical decoupling in the TAT protocol}
+
+The effective Hamiltonian resulting from a perturbative treatment of SOC is (see Methods in the main text)
+\begin{align}
+  H_{\t{eff}}
+  = -\f{U}{L}\v S\c\v S - \overline B S_\z + \Omega S_\x + \chi S_\z^2,
+  \label{eq:H_eff_DD}
+\end{align}
+where $U$ is a two-atom on-site interaction strength; $L$ is the number of lattice sites; $\overline B\equiv \sum_n B_n/N$ is a residual axial field determined by the occupied quai-momentum modes $\set{n}$ (with $\abs{\set{n}}=N$ atoms total); $\Omega$ is the magnitude of a driving field; and $\chi$ is an effective OAT squeezing strength.
+The effect of the $\sim\v S\c\v S$ term is to generate a relative phase between states with different total spin $S$ (where $S=N/2$ within the Dicke manifold).
+In the absence of coherent coupling between states with different total spin, therefore, the $\sim\v S\c\v S$ term has no effect on system dynamics, and we are safe to neglect it entirely.
+
+In the parameter regimes relevant to our discussions in the main text, the operator norms of $\overline{B}\hat S_\z$ and $\chi \hat S_\z^2$ in Eqn.~\eqref{eq:H_eff_DD} will typically be comparable in magnitude.
+The OAT protocol sets $\Omega=0$, and eliminates the effect of $\overline{B}\hat S_\z$ with a spin-echo $\pi$-pulse $\exp\p{-i\pi\hat S_\x}$ applied half way through the squeezing protocol.
+The TAT protocol, meanwhile, effectively takes $\chi\hat S_\z^2+\Omega \hat S_\x\to\hat H_{\t{TAT}}^{(\pm)}$ (as defined in the Methods of the main text) and $\overline{B}\hat S_\z\to\J_0\p{\beta_\pm}\overline{B}\hat S_\z$, where $\J_0$ is the zero-order Bessel function of the first kind and $\beta_\pm$ is the modulation index of the amplitude-modulated driving field $\Omega$, satisfying $\J_0\p{2\beta_\pm}=\pm1/3$.
+Unlike in the case of OAT, $\hat S_\z$ does not commute with the TAT Hamiltonian, so its effect cannot be eliminated with a spin-echo.
+Nonetheless, this term can be eliminated with a dynamical decoupling pulse sequence that periodically inverts the sign of $\hat S_\z$ while preserving $\hat H_{\t{TAT}}^{(\pm)}$.
+
+\begin{spacing}{1}
+\begin{figure}
+\centering
+\includegraphics{pulsed_squeezing.pdf}
+\caption{{\bf Optimal squeezing as a function of $\pi$-pulses} applied prior to the optimal TAT squeezing time in a CPMG sequence with ({\bf a)} $N=100$ and ({\bf b}) $N=1000$ atoms.
+Results are shown for OAT, TAT, and TAT$_{\pm,\z}$, where TAT$_{\pm,\z}$ denotes squeezing via the Hamiltonian $\hat H_{\t{TAT}}^{(\pm,\z)}\equiv \hat H_{\t{TAT}}^{(\pm)}-\J_0\p{\beta_\pm}\bk{\overline B}^{\t{rms}}_f \hat S_\z$.
+Details about experimental parameters for these simulations are provided in the text.
+}
+\label{fig:pulsed_squeezing}
+\end{figure}
+\end{spacing}
+
+Fig.~\ref{fig:pulsed_squeezing} shows the maximal squeezing generated by $N=10^2$ and $10^3$ atoms via OAT, TAT, and TAT in the presence of the mean field $\J_0\p{\beta_\pm}\overline B\hat S_\z$ as a function of the number of $\pi$-pulses performed prior to the optimal TAT squeezing time.
+These pulses are applied in a CPMG sequence $\p{\tau_n/2-\pi_\x-\tau_n/2}^n$, where $\tau_n/2$ denotes Hamiltonian evolution for a time $\tau_n/2$, $\pi_\x$ denotes the application of an instantaneous $\pi$-pulse $\exp\p{-i\pi\hat S_\x}$, and $n$ is the number of pulses, such that the optimal TAT squeezing time is $t_{\t{opt}}^{\t{TAT}}=\p{\tau_n}^n$.
+The label TAT$_{\pm,\z}$ in Fig.~\ref{fig:pulsed_squeezing} denotes squeezing through the Hamiltonian $\hat H_{\t{TAT}}^{(\pm,\z)}\equiv \hat H_{\t{TAT}}^{(\pm)}-\J_0\p{\beta_\pm}\bk{\overline B}^{\t{rms}}_f\hat S_\z$, where $\bk{\overline B}^{\t{rms}}_f$ is the root-mean-square average of $\overline B$ over choices of occupied spacial modes $\set{n}$ at fixed filling $f$ of all spatial modes in the lowest Bloch band of a periodic 2D lattice.
+While the modulation index $\beta_+$ is uniquely defined by $\J_0\p{2\beta_+}=1/3$, there are two choices of $\beta_-$ for which $\J_0\p{2\beta_-}=-1/3$; we use that which minimizes $\abs{\J_0\p{\beta_-}}$.
+Fig.~\ref{fig:pulsed_squeezing} assumes an SOC angle $\phi=\pi/50$ (although results are independent of $\phi$ for $\phi\ll1$), a reduced field variance $\widetilde{B}/U=0.05$, and a filling $f=5/6$.
+Note that as the filling $f\to1$, the residual axial field vanishes ($\overline{B}\to0$), so TAT$_{\pm,\z}\to$ TAT.
+
+\section*{Decoherence in the 3D $^{87}$Sr optical lattice clock}
+
+Currently, light scattering from lattice beams in the 3D $^{87}$Sr optical lattice clock induces decoherence on a time scale of $\sim$10 seconds\cite{goban2018emergence, hutson2019engineering}.
+This decoherence acts identically on all atoms in an uncorrelated manner, and can be understood by considering the density operator $\rho$ for a single atom, with effective spin states $\dn$ and $\up$ respectively corresponding to the ${}^1\t{S}_0$ and ${}^3\t{P}_0$ electronic states.
+Empirically, the effect of decoherence after a time $t$ within the $\set{\dn,\up}$ subspace of a single atom is to take $\rho\to\rho\p{t}$ with $\rho\p{0}\equiv\rho$ and
+\begin{align}
+  \rho\p{t} \coloneqq
+  \begin{pmatrix}
+  \rho_{\up\up} e^{-\Gamma_{\up\up}t} &&
+  \rho_{\up\dn} e^{-\Gamma_{\up\dn}t} \\
+  \rho_{\up\dn}^* e^{-\Gamma_{\up\dn}t} &&
+  \rho_{\dn\dn} + \p{1-e^{-\Gamma_{\up\up}t}} \rho_{\up\up}
+  \end{pmatrix},
+  \label{eq:decay_matrix}
+\end{align}
+where $\Gamma_{\up\up}\approx\Gamma_{\up\dn}\approx\Gamma=0.1~\t{sec}^{-1}$ are respectively decay rates for $\rho_{\up\up}$ and $\rho_{\up\dn}$.
+This form of decoherence can be effectively modeled by decay and dephasing of individual spins (respectively denoted $\Gamma_{\t{ud}}$ and $\Gamma_{\t{el}}$ in Ref.~[\citenum{foss-feig2013nonequilibrium}]) at rates $\Gamma$.
+In the language of the section that follows, we would say that this decoherence is captured by the sets of jump operators $\J_-\equiv\set{s_-^{(j)}}$ and $\J_\z\equiv\set{s_\z^{(j)}}$ with corresponding decoherence rates $\gamma_-=\gamma_\z=\Gamma$.
+
+\section*{Time-series of squeezing via OAT and TAT}
 
@@
 In the absence of decoherence, OAT achieves a maximal amount of squeezing that scales as $\xi^2\sim N^{-2/3}$, while TAT achieves Heisenberg-limited squeezing with $\xi^2\sim N^{-1}$.
-Note that our method for computing squeezing via TAT in the presence of decoherence (described in Appendix \ref{M-sec:collective_simulation} of the main text) is not capable of computing squeezing for the full range of times shown in Fig.~\ref{fig:squeezing_example}; the corresponding time-series data in this figure is therefore shown up to the point at which this method breaks down.
+Note that our method for computing squeezing via TAT in the presence of decoherence (described in the Methods of the main text) is not capable of computing squeezing for the full range of times shown in Fig.~\ref{fig:squeezing_example}; the corresponding time-series data in this figure is therefore shown up to the point at which this method breaks down.
 
-\begin{figure*}
+\begin{spacing}{1}
+\begin{figure}
 \centering
@@
 Atoms are confined along the direction transverse to the 2D layer by a lattice of depth 60 $E_{\t{R}}$.
-Squeezing over time is shown for OAT (blue) and TAT (green), both with (solid lines) and without (dashed lines) decoherence via uncorrelated decay and dephasing of individual spins at rates of $0.1~\t{sec}^{-1}$ (see Appendix \ref{M-sec:decoherence} of the main text).
+Squeezing over time is shown for OAT (blue) and TAT (green), both with (solid lines) and without (dashed lines) decoherence via uncorrelated decay and dephasing of individual spins at rates of $0.1~\t{sec}^{-1}$ (see the section on decoherence in the 3D clock above).
 }
 \label{fig:squeezing_example}
-\end{figure*}
+\end{figure}
+\end{spacing}
+
+\section*{Solving Heisenberg equations of motion for collective spin systems}
+
+In order to compute squeezing of a collective spin system, we need to compute expectation values of (homogeneous) collective spin operators.
+We compute these expectation values using a method recently developed in Ref.~[\citenum{perlin2019shorttime}], and provide a short overview of the method here.
+Choosing the basis $\set{\S_{\v m}}$ for all collective spin operators, where $\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$ with
+$\v m\equiv\p{m_+,m_\z,m_-}\in\mathbb{N}_0^3$, we can expand all
+collective spin Hamiltonians in the form
+\begin{align}
+  H = \sum_{\v m} h_{\v m} \S_{\v m}.
+  \label{eq:H_general}
+\end{align}
+The evolution of a general correlator $\bk{\S_{\v n}}$ under a Hamiltonian of the form in Eqn.~\eqref{eq:H_general} is then given by
+\begin{align}
+  \f{d}{dt} \bk{\S_{\v n}}
+  = i \sum_{\v m} h_{\v m}
+  \bk{\sp{\S_{\v m}, \S_{\v n}}_-}
+  + \sum_\J \gamma_\J \bk{\mathcal D\p{\J; \S_{\v n}}}
+  \equiv \sum_{\v m} \bk{\S_{\v m}} T_{\v m\v n},
+\end{align}
+where $\sp{X,Y}_\pm\equiv XY\pm YX$; $\J$ is a set of jump operators with corresponding decoherence rate $\gamma_\J$; the decoherence operator $\mathcal D$ is defined by
+\begin{align}
+  \mathcal{D}\p{\J;\O}
+  \equiv \sum_{J\in\J}\p{J^\dag \O J - \f12\sp{J^\dag J,\O}_+};
+\end{align}
+and $T_{\v m\v n}$ is a matrix element of the time derivative operator $T\equiv d/dt$.
+These matrix elements can be calculated analytically using product and commutation rules for collective spin operators.
+We can then expand correlators in a Taylor series about $t=0$ to write
+\begin{align}
+  \bk{\S_{\v n}}
+  &= \sum_{k\ge0} \f{t^k}{k!} \bk{\f{d^k}{dt^k} \S_{\v n}}_{t=0}
+  = \sum_{k\ge0} \f{t^k}{k!}
+  \sum_{\v m} T_{\v m\v n;k} \bk{\S_{\v m}}_{t=0},
+  \label{eq:time_series}
+\end{align}
+where $T_{\v m\v n;k}\equiv\sp{T^k}_{\v m\v n}$ are matrix elements of the $k$-th time derivative.
+Expectation values of collective spin operators can thus be computed via the expansion in Eqn.~\eqref{eq:time_series}, which at short times can be truncated at some finite order beyond which all terms have negligible contribution to $\bk{\S_{\v n}}$.
+
+\section*{Accounting for $p$-wave inelastic collisions}
+
+\begin{spacing}{1}
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.4\textwidth]{rates.pdf}
+\caption{
+Both the averaged $p$-wave inelastic collision rate $\gamma$ (orange) and the ratio of this collision rate to the optimal squeezing rate $\chi_{\text{opt}}$ (blue) are suppressed as the lattice depth increases.
+%The blue line shows  the ratio of the inelastic collision rate $\gamma$ and the optimal squeezing rate $\chi_{\text{opt}}$.
+$\chi_{\text{opt}}$ is obtained by choosing SOC angles $\phi$ that saturate $\widetilde{B}/U\approx0.05$, where $\widetilde{B}$ is the variance of the SOC-induced axial field and $U$ is the two-atom on-site interaction energy.
+%The ratio decreases with increasing lattice depth.
+}
+\label{fig:inelastic_rates}
+\end{figure}
+\end{spacing}
+
+\begin{spacing}{1}
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.7\textwidth]{inelasticPlot.pdf}
+\caption{\textbf{Squeezing via OAT in the presence of inelastic collisions}.
+(a) For fixed particle number $N=100$, the optimal squeezing decreases as the inelastic collision rate increases.
+Panel (b) shows squeezing over time for $\gamma/\chi_{\text{opt}}=0.04$ (solid lines), which corresponds to $U/J=6$,
+and compares it with $\gamma=0$ (dashed lines) for  different particle numbers.
+Inelastic collisions prevent the growth of optimal  squeezing with  particle number. For $N=1000$, the maximum squeezing saturates to $\sim 10$dB.
+}
+\label{fig:inelastic_squeezing}
+\end{figure}
+\end{spacing}
+
+Inelastic $^{3}\text{P}_{0}$ (electronic state $e$ or $\up$) collisions are detrimental for optical lattice clocks.
+For the  nuclear-spin-polarized gas discussed in this work, $ee$ losses are only possible via the $p$-wave scattering channel since $s$-wave collisions are suppressed by Fermi statistics.
+The big advantage here compared to  prior experiments done in a 1D lattice at $\mu$K temperature\cite{martin2013quantum} is that in a Fermi  degenerate  gas loaded in a 3D optical lattice, $p$-wave losses are further suppressed by the centrifugal barrier and Pauli blocking, and only happen through a wave-function overlap between atoms at different lattice sites.
+In this appendix, we quantify the effect of $p$-wave interactions on squeezing.
+To account for $p$-wave losses, we describe the dynamics  using a master equation for the system's density matrix $\hat{\rho}$:
+\begin{align}
+  \frac{d\hat{\rho}}{dt}
+  = -i [\hat{H}_{\text{eff}},\,\hat{\rho}]
+  + \mathcal{L}\hat{\rho},
+\end{align}
+where $\hat{H}_{\text{eff}}=\chi\hat{S}_z^2$ is the effective one-axis twisting Hamiltonian obtained from  the original Fermi-Hubbard Hamiltonian with spin-orbit coupling, and $\mathcal{L}$ is the Lindblad superoperator that accounts for $p$-wave $ee$ inelastic collisions.
+This latter term can be written using a pseudo-potential approximation as\cite{rey2014probing}:
+\begin{align}
+  \mathcal{L}\hat\rho
+  = \sum_{{\bm k}{\bm k'}} \Gamma_{{\bm k}{\bm k'}}
+  \sp{\hat{A}_{{\bm k}{\bm k'}} \hat\rho\hat{A}_{{\bm k}{\bm k'}}^\dag
+  - \f12 \p{\hat{A}_{{\bm k}{\bm k'}}^\dag \hat{A}_{{\bm k}{\bm k'}} \hat\rho
++\hat\rho\hat{A}_{{\bm k}{\bm k'}}^\dag \hat{A}_{{\bm k}{\bm k'}}}},
+\end{align}
+where the jump operators are $\hat{A}_{{\bm k}{\bm k'}}=\hat{c}_{{\bm k},\up}\hat{c}_{{\bm k'},\up}$, and
+$\bm k$, $\bm k'$ sum over all the populated quasi-momentum modes. The decay matrix elements $\Gamma_{{\bm k}{\bm k'}}$ are given by:
+\begin{align}
+  \Gamma_{{\bm k}{\bm k'}}
+  = \frac{3\pi b_{\text{im}}^3 }{m} \p{\int d{\bm r}^{\,3}
+    W[\phi_{{\bm k}}({\bm r}),\phi_{{\bm k'}}({\bm r})]},
+\end{align}
+where $b_{\text{im}} = 121 a_0$\cite{zhang2014spectroscopic, goban2018emergence} is the $p$-wave inelastic scattering length (with $ a_0=5.29\times 10 ^{-11}$ m the Bohr radius), $\phi_{{\bm k}}({\bm r})$ is the Bloch function with quasi-momentum ${\bm k}$, and
+\begin{multline}
+  W\sp{\phi_{{\bm k}}({\bm r}),\phi_{{\bm k'}}({\bm r})} \\
+  \equiv \sp{\p{{\bm \nabla}\phi^{*}_{{\bm k}}(\bm r)}
+  \phi^{*}_{{\bm k'}}(\bm r)-\phi^{*}_{{\bm k}}(\bm r)
+  \p{{\bm \nabla}\phi^{*}_{{\bm k'}}(\bm r)}}
+  \cdot\sp{\p{{\bm \nabla}\phi_{{\bm k}}(\bm r)}
+  \phi_{{\bm k'}}(\bm r)-\phi_{{\bm k}}(\bm r)
+  \p{{\bm \nabla}\phi_{{\bm k'}}(\bm r)}}
+\end{multline}
+
+In Fig.~\ref{fig:inelastic_rates} we show  the averaged decay rate $\gamma\equiv\sum_{{\bm k}{\bm k'}}\Gamma_{{\bm k}{\bm k'}}/\ell^2$, where $\ell$ is the number of lattice sites along the $x$ and $y$ axes, as a function of the  lattice depth $V_0$ along these axes.
+Here we assume  the same lattice depth  in the $z$ direction used in the main text, $V=60E_{\text{R}}$.
+As can be seen, the decay rate $\gamma$ is largely suppressed as $V_0$ increases  but is not completely negligible at shallow lattices.
+To quantify the effect of these losses on the spin squeezing generation process, we follow a similar methodology to the one described in detail in Ref.~\cite{rey2014probing}.
+The basic idea is to take advantage of the so-called Truncated-Wigner Approximation (TWA)\cite{polkovnikov2010phase, schachenmayer2015manybody}, which allows us to capture the development of spin squeezing using semi-classical phase-space methods.
+In the  TWA the quantum dynamics are accounted for by solving mean field equations of motion supplemented by noise.
+The mean field equations are  derived by assuming that the many-body density matrix of the system can be factorized as
+$\hat{\rho}= \bigotimes_{i} \hat{\rho}({i})$, where $\hat{\rho}(i)$ is the reduced density matrix of the  particle in quasi-momentum  mode  ${\bm q}_i$ [see Eqn.~\eqref{eq:decay_matrix}].
+Under this assumption, the non-linear mean field equations are given by
+\begin{align}
+  \frac{d\rho_{\up\up}(j)}{dt}
+  = -\sum_{j'}\Gamma_{{\bm k}_j{\bm k}_{j'}}
+  \rho_{\up\up}(j) \rho_{\up\up}(j'),
+  &&
+  \frac{d\rho_{\dn \dn}(j)}{dt}=0
+\end{align}
+\begin{align}
+  \frac{d\rho_{\up\dn}(j)}{dt}
+  = \rho_{\up\dn}(j) \sum_{j'}
+  \sp{i\chi(\rho_{\up\up}(j')-\rho_{\dn\dn}(j'))
+  - \f12 \Gamma_{{\bm k}_j{\bm k}_{j'}} \rho_{\up\up}(j')},
+\end{align}
+where $\rho_{\sigma\sigma'}\equiv\langle\hat{\rho}_{\sigma\sigma'}\rangle$.
+Since we are interested in the collective behavior, one can define $\rho_{\sigma\sigma'}^T=\sum_j \rho_{\sigma\sigma'}(j)$. For these observables the equations of motion simplify to
+\begin{align}
+  \frac{d\rho_{\up\up}^T}{dt} = -f\gamma  (\rho_{\up\up}^T)^2,
+  && \frac{d\rho_{\dn \dn}^T}{dt} = 0
+  &&
+  \frac{d\rho_{\up\dn}^T}{dt}
+  = \rho_{\up\dn}^T
+  \sp{i\chi(\rho_{\up\up}^T-\rho_{\dn\dn}^T)
+    - \f12 f\gamma \rho_{\up\up}^T},
+\end{align}
+where $f\equiv N/\ell^2$ is the filling fraction.
+
+Under the TWA, one accounts for quantum fluctuations during the dynamics by  averaging over different mean field trajectories generated by sampling over different initial conditions chosen to reconstruct the Wigner function of the initial coherent spin state at $t=0$\cite{rey2014probing}.
+This method has proven to be successful in simulating quantum spin dynamics.
+Using this approach, Fig.~\ref{fig:inelastic_squeezing} shows numerical simulation results of squeezing over time in the presence of inelastic collisions.
+For shallow lattices, the effect of inelastic collisions can limit the spin squeezing to 10 dB.
+Thus, in this regime, losses are as relevant as light scattering.
+The role of inelastic interactions could be mitigated by either operating at deeper lattices as shown in Fig.
+\ref{fig:inelastic_squeezing}, or by using nuclear spin states to generate the squeezing instead of the clock states directly.
+
+\bibliography{main.bib}
 

### 2019-09-05 soc_squeezing/collective_paper/collective_spin.tex
some changes to collective spin paper from internal review

@@
 \newcommand{\bk}{\braket} % shorthand for braket notation
+\newcommand{\Bk}{\Braket}
 
@@
 \newcommand{\D}{\mathcal{D}}
-\newcommand{\F}{\mathcal{F}}
-\newcommand{\I}{\mathcal{I}}
+\newcommand{\E}{\mathcal{E}}
+\newcommand{\J}{\mathcal{J}}
 \renewcommand{\O}{\mathcal{O}}
@@
 \usepackage[inline]{enumitem} % for inline enumeration
+\setenumerate{label=(\roman*)}
 
@@
   klinder2015dynamical}, ultracold Fermi
-gasses\cite{martin2013quantum, smale2018observation}, trapped
+gasses\cite{martin2013quantum, smale2019observation}, trapped
 ions\cite{bohnet2016quantum}, optical
@@
   norcia2018cavitymediated}, and molecular
-sensors\cite{jones2009magnetic}.  These implementations have triggered
-studies of a variety of rich subjects, including quantum
-criticality\cite{latorre2005entanglement, alcalde2007functional,
-  morrison2008dissipationdriven, morrison2008collective,
-  sarandy2009classical, wang2012quantum, kessler2012dissipative,
-  majd2014lmg}, non-equilibrium
+sensors\cite{jones2009magnetic}, among others.  These implementations
+have triggered studies of a variety of rich subjects, including
+quantum criticality\cite{latorre2005entanglement,
+  alcalde2007functional, morrison2008dissipationdriven,
+  morrison2008collective, sarandy2009classical, wang2012quantum,
+  kessler2012dissipative, majd2014lmg}, non-equilibrium
 phenomena\cite{walls1978nonequilibrium, morrison2008dynamical,
@@
   maghrebi2016nonequilibrium, zhiqiang2017nonequilibrium,
-  lang2018concurrence, smale2018observation}, and precision
+  lang2018concurrence, smale2019observation}, and precision
 metrology\cite{wineland1992spin, itano1993quantum,
@@
 short-time expansion.  Finally, we showcase applications of our method
-with computations that are inaccessible to other numerical methods.
+by computing quantities that are inaccessible to other numerical
+methods.
 
@@
 self-adjoint, for example, then $\O_{\v m}^*=\O_{\v m^*}$ with
-$\v m^*\equiv\p{m_-,m_\z,m_+}$.  The corresponding Heisenberg operator
-is then $\hat\O\p{t}=\sum_{\v m}\O_{\v m}\p{t}\hat\S_{\v m}$, with
-time-dependent coefficients $\O_{\v n}\p{t}$ for time-independent
-(Schr\"odinger) operators $\hat\S_{\v m}$.  The time evolution of
-Heisenberg operators is determined by
-\begin{align}
-  \f{d}{dt} \hat\O\p{t}
-  = \check T \hat\O\p{t}
-  = \sum_{\v m, \v n} \hat\S_{\v m} T_{\v m\v n} \O_{\v n}\p{t}
+$\v m^*\equiv\p{m_-,m_\z,m_+}$.
+%
+% The corresponding Heisenberg operator is then
+% $\hat\O\p{t}=\sum_{\v m}\O_{\v m}\p{t}\hat\S_{\v m}$, with
+% time-dependent coefficients $\O_{\v n}\p{t}$ for time-independent
+% (Schr\"odinger) operators $\hat\S_{\v m}$.  The time evolution of
+% Heisenberg operators is determined by
+%
+The corresponding Heisenberg operator is then
+$\hat\O\p{t}=\sum_{\v m}\O_{\v m}\p{t}\hat\S_{\v m}+\hat\E_\O\p{t}$,
+with time-dependent coefficients $\O_{\v n}\p{t}$ for time-independent
+(Schr\"odinger) operators $\hat\S_{\v m}$, and mean-zero ``noise''
+operators $\hat\E_\O\p{t}$ that result from interactions between the
+spin system and its environment, initially $\hat\E_\O\p{0}=0$.  The
+expectation values of Heisenberg operators evolve according to
+%
+\begin{align}
+  \f{d}{dt} \bk{\hat\O\p{t}}
+  = \bk{\check T \hat\O\p{t}}
+  = \sum_{\v m, \v n} \bk{\hat\S_{\v m}} T_{\v m\v n} \O_{\v n}\p{t}
   \label{eq:EOM}
 \end{align}
-for a time derivative (super-)operator $\check T$ with matrix elements
-$T_{\v m\v n}\in\C$ defined by
+with a time derivative (super-)operator $\check T$ whose matrix
+elements $T_{\v m\v n}\in\C$ are defined by
 \begin{align}
   \check T \hat\S_{\v n} \equiv i \sp{\hat H, \hat\S_{\v n}}_-
-  + \sum_\A \gamma_\A \check\D\p{\A} \hat\S_{\v n}
+  + \sum_\J \gamma_\J \check\D\p{\J} \hat\S_{\v n}
   = \sum_{\v m} \hat\S_{\v m} T_{\v m\v n},
@@
 where $\sp{X,Y}_\pm\equiv XY\pm YX$; $\hat H$ is the collective spin
-Hamiltonian; $\A$ is a set of jump operators with a corresponding
-decoherence rate $\gamma_\A$; and $\check\D$ is a Heisenberg-picture
+Hamiltonian; $\J$ is a set of jump operators with a corresponding
+decoherence rate $\gamma_\J$; and $\check\D$ is a Heisenberg-picture
 dissipator, or Lindblad superoperator, defined by
 \begin{align}
-  \check\D\p{\A} \hat \O
-  \equiv \sum_{\hat A\in\A} \p{\hat A^\dag \hat \O \hat A
-    - \f12\sp{\hat A^\dag \hat A, \hat \O}_+}.
+  \check\D\p{\J} \hat \O
+  \equiv \sum_{\hat J\in\J} \p{\hat J^\dag \hat \O \hat J
+    - \f12\sp{\hat J^\dag \hat J, \hat \O}_+}.
 \end{align}
@@
 would be described by the set of jump operators
-$\A_-\equiv\set{\hat s_-^{(j)}:j=1,2,\cdots,N}$.  The commutator in
+$\J_-\equiv\set{\hat s_-^{(j)}:j=1,2,\cdots,N}$.  The commutator in
 Eq.~\eqref{eq:time_deriv} can be computed by expanding the product
@@
 \ref{sec:identities}--\ref{sec:general_product}, and the effects of
-decoherence from jump operators (i.e.~elements of $\A$) of the form
+decoherence from jump operators (i.e.~elements of $\J$) of the form
 $\hat g^{(j)} = \sum_\a g_\a \hat s_\a^{(j)}$ and
@@
 \begin{align}
-  \f{d}{dt} \hat\O\p{t}
-  \to \sum_{w\p{\v m}<W} \hat\S_{\v m}
+  \f{d}{dt} \bk{\hat\O\p{t}}
+  \to \sum_{w\p{\v m}<W} \bk{\hat\S_{\v m}}
   \sum_{\v n} T_{\v m\v n} \O_{\v n}\p{t}
@@
 limitations:
-\begin{enumerate*}[label=(\roman*)]
+\begin{enumerate*}
 \item simulating a system of differential equations for a large number
@@
 \begin{align}
-  \hat\O\p{t}
-  = e^{t\check T} \hat \O\p{0}
+  \bk{\hat\O\p{t}}
+  = \bk{e^{t\check T} \hat \O\p{0}}
   = \sum_{k\ge0} \f{t^k}{k!}
-  \sum_{\v m, \v n} \hat\S_{\v m} T^k_{\v m\v n} \O_{\v n}\p{0},
+  \sum_{\v m, \v n} \bk{\hat\S_{\v m}} T^k_{\v m\v n} \O_{\v n}\p{0},
   \label{eq:time_series}
@@
 where the matrix elements $T^k_{\v m\v n}$ of the $k$-th time
-derivative operator $T^k$ are
+derivative operator $\check T^k$ are
 \begin{align}
@@
 \begin{align}
-  \hat\O\p{t}
+  \bk{\hat\O\p{t}}
   \to \sum_{k=0}^M \f{t^k}{k!}
-  \sum_{\v m, \v n} \hat\S_{\v m} T^k_{\v m\v n} \O_{\v n}\p{0}.
+  \sum_{\v m, \v n} \bk{\hat\S_{\v m}} T^k_{\v m\v n} \O_{\v n}\p{0}.
   \label{eq:TST}
 \end{align}
+%
+% We refer Eq.~\eqref{eq:TST} as the truncated short-time (TST)
+% expansion of Heisenberg operators.  Unlike the truncation in
+% Eq.~\eqref{eq:weight_truncation}, the nonzero matrix elements
+% $T^k_{\v m\v n}$ for $k=0,1,\cdots,M$ in Eq.~\eqref{eq:TST} tell us
+% which operators are relevant for computing the expectation value
+% $\bk{\hat\O\p{t}}$ to a fixed order $M$.  Note that using the
+% relation $\hat\S_{\v m}^\dag=\hat\S_{\v m^*}$, which by Hermitian
+% conjugation of Eq.~\eqref{eq:EOM} also implies that
+% $T_{\v m^*\v n^*}=T_{\v m\v n}^*$, can cut both the number of
+% initial-time expectation values $\bk{\hat\S_{\v m}}_{t=0}$ and the
+% number of matrix elements $T_{\v m\v n}$ that we may need to
+% explicitly compute roughly in half.
+%
 We refer Eq.~\eqref{eq:TST} as the truncated short-time (TST)
-expansion of Heisenberg operators.  Unlike the truncation in
+expansion of Heisenberg operators.  Note that when computing an
+expectation value $\bk{\hat\O\p{t}}$, the relation
+$\hat\S_{\v m}^\dag=\hat\S_{\v m^*}$, which by Hermitian conjugation
+of Eq.~\eqref{eq:EOM} also implies that
+$T_{\v m^*\v n^*}=T_{\v m\v n}^*$, cuts both the number of
+initial-time expectation values $\bk{\hat\S_{\v m}}$ and the number of
+matrix elements $T_{\v m\v n}$ that we may need to explicitly compute
+roughly in half.
+
+Unlike the weight-based truncation in
 Eq.~\eqref{eq:weight_truncation}, the nonzero matrix elements
 $T^k_{\v m\v n}$ for $k=0,1,\cdots,M$ in Eq.~\eqref{eq:TST} tell us
-which operators are relevant for computing the expectation value
-$\bk{\hat\O\p{t}}$ to a fixed order $M$.  Note that using the relation
-$\hat\S_{\v m}^\dag=\hat\S_{\v m^*}$, which by Hermitian conjugation
-of Eq.~\eqref{eq:EOM} also implies that
-$T_{\v m^*\v n^*}=T_{\v m\v n}^*$, can cut both the number of
-initial-time expectation values $\bk{\hat\S_{\v m}}_{t=0}$ and the
-number of matrix elements $T_{\v m\v n}$ that we may need to
-explicitly compute roughly in half.
+which operators $\hat\S_{\v m}$ are relevant for computing the
+expectation value $\bk{\hat\O\p{t}}$ to a fixed order $M$.  The TST
+expansion thereby avoids the introduction of a weight measure $w$ that
+chooses which operators to keep track of, and trades the cost of
+solving a system of differential equations for the cost of computing
+expectation values $\bk{\hat\S_{\v m}}$ and matrix elements
+$T_{\v m\v n}^k$.  In all cases considered in this work, we find that
+the TST expansion is both faster to evaluate and provides accurate
+correlators $\bk{\hat\O\p{t}}$ until later times $t$ than the
+weight-based expansion in \eqref{eq:weight_truncation} with weight
+measure $w\p{\v m}=\abs{\v m}$ and cutoff $W\approx M$.  We therefore
+restrict the remainder of our discussions to the TST expansion in
+Eq.~\eqref{eq:TST}.
+%
 
@@
 computational resources, such that the TST expansion in
-Eq.~\eqref{eq:TST} only allows for computation of the expectation
-value $\bk{\hat\O\p{t}}$ to a maximal time $t_{M_{\t{max}}}$.  As we
-will see in the following section, $M_{\t{max}}=35$ will suffice for
-the computation of collective spin correlators up to times that are
+Eq.~\eqref{eq:TST} only allows us to compute the expectation value
+$\bk{\hat\O\p{t}}$ to a maximal time $t_{M_{\t{max}}}$.  As we will
+see in the following section, $M_{\t{max}}=35$ will suffice for the
+computation of collective spin correlators up to times that are
 relevant e.g.~for spin squeezing protocols.  Appendix
@@
 TNT in the large-$N$ limit\cite{micheli2003manyparticle,
-  sorelli2015fast}.
+  sorelli2019fast}.
 
@@
   \min_{\substack{\v v\perp\bk{\uv S}\\\v v\cdot\v v=1}}
-    \Braket{\p{\uv S\cdot\v v}^2},
+  \Bk{\p{\uv S\cdot\v v}^2},
   \label{eq:squeezing}
@@
 \ref{fig:squeezing_incoherent}), respectively described by the sets of
-jump operators $\A_\a\equiv\set{\hat s_\a^{(j)}}$ with corresponding
+jump operators $\J_\a\equiv\set{\hat s_\a^{(j)}}$ with corresponding
 decoherence rates $\gamma_\a=\chi$ for $\a\in\set{-,+,\z}$.  The
@@
 individual correlators taking unphysical values
-$\lvert\bk{\hat\S_{\v n}}\rvert\gtrsim S^{\abs{\v n}}$.  The sudden
-and drastic departure from virtually exact results occurs because the
-TST expansion neglects high-weight (i.e.~large-$\abs{\v m}$) operators
-$\hat\S_{\v m}$ whose contributions to a Heisenberg operator
-$\hat\O\p{t}$ of interest eventually become non-negligible.  For
-sufficiently large times $t$, these individual contributions may
-generally be large compared to the value of $\bk{\hat\O\p{t}}$, in
-which case their truncation will neglect terms that are large compared
-to the actual value of $\bk{\hat\O\p{t}}$.  This breakdown mechanism
+$\lvert\bk{\hat\S_{\v n}\p{t}}\rvert\gtrsim S^{\abs{\v n}}$.  The
+sudden and drastic departure from virtually exact results occurs
+because the TST expansion neglects high-weight
+(i.e.~large-$\abs{\v m}$) operators $\hat\S_{\v m}$ whose
+contributions to a Heisenberg operator $\hat\O\p{t}$ of interest
+eventually become non-negligible.  For sufficiently large times $t$,
+these individual contributions can become comparable in magnitude to
+the expectation value $\bk{\hat\O\p{t}}$, and will grow with some
+large power of $t$, so their truncation will have a drastic effect on
+the estimated value of $\bk{\hat\O\p{t}}$.  This breakdown mechanism
 is also the reason for prolonged agreement between the TST expansion
@@
 rather than $\ket{\X}$: for initial states $\ket{-\Z}$, all
-initial-time correlators $\bk{\hat\S_{\v m}}_{t=0}$ vanish unless
+initial-time correlators $\bk{\hat\S_{\v m}}$ vanish unless
 $m_+=m_-=0$ (see Appendix \ref{sec:initial_conditions}), so there is a
@@
 contribution to the correlator $\bk{\hat\O\p{t}}$.
-
-Although the TST expansion breaks down at short times, it has two key
-advantages over other methods to compute collective spin correlators.
-First, computing spin correlators with the TST expansion is generally
-much faster and requires much fewer computing resources than the
-alternatives.  The quantum trajectory Monte Carlo simulations
-performed for Figure \ref{fig:squeezing_incoherent}, for example, take
-$\sim10^4$ CPU hours to compute on standard modern computing hardware;
-the bulk of this time is spent performing sparse matrix-vector
-multiplication, leaving little room to further optimize run time.
-Parallelization can reduce the actual run time of these quantum
-trajectory simulations to $\sim10$ hours, but at the cost of greatly
-increasing computing resource requirements.  The TST expansion results
-in Figure \ref{fig:squeezing_incoherent}, meanwhile, take $\sim10$
-seconds to compute with a single CPU on similar hardware.
 
@@
 
+%
+% Although the TST expansion breaks down at short times, it has two
+% key advantages over other methods to compute collective spin
+% correlators.  First, computing spin correlators with the TST
+% expansion is generally much faster and requires much fewer computing
+% resources than the alternatives.  The quantum trajectory Monte Carlo
+% simulations performed for Figure \ref{fig:squeezing_incoherent}, for
+% example, take $\sim10^4$ CPU hours to compute on standard modern
+% computing hardware; the bulk of this time is spent performing sparse
+% matrix-vector multiplication, leaving little room to further
+% optimize serial runtime.  Parallelization can reduce the actual
+% runtime of these quantum trajectory simulations to $\sim10$ hours by
+% running all trajectories at once, but at the cost of greatly
+% increasing computing resource requirements.  The TST expansion
+% results in Figure \ref{fig:squeezing_incoherent}, meanwhile, take
+% $\sim10$ seconds to compute with a single CPU on similar hardware.
+%
+Although the TST expansion breaks down at short times, it has two key
+advantages over the quantum trajectory Monte Carlo method to compute
+correlators in the presence of decoherence.  First, computing spin
+correlators with the TST expansion is generally faster and requires
+fewer computing resources.  The TST expansion results in Figure
+\ref{fig:squeezing_incoherent}, for example, take $\sim10$ seconds to
+compute with a single CPU on modern computing hardware.  The quantum
+trajectory Monte Carlo results in the same figure, meanwhile, take
+$\sim10^4$ CPU hours to compute on similar hardware; the bulk of this
+time is spent performing sparse matrix-vector multiplication, leaving
+little room to further optimize serial runtime.  Parallelization can
+reduce actual runtime of the Monte Carlo simulations to $\sim10$ hours
+by running all trajectories at once, but at the cost of greatly
+increasing computing resource requirements.  Though it may be possible
+to further speed up quantum trajectory Monte Carlo simulations by
+introducing new truncation schemes, any modifications
+\begin{enumerate*}
+\item should be made carefully to ensure that simulations still yield
+  correct results, and
+\item are unlikely to bridge the orders of magnitude in computing
+  resource requirements.
+\end{enumerate*}
+%
+
 Second, the TST expansion enables computing spin correlators in
-strong-decoherence regimes entirely inaccessible to other methods.  As
-an example, Figure \ref{fig:decoherence_strong} shows squeezing of
-$N=10^4$ spins initially in $\ket{\X}$, undergoing spontaneous decay,
-excitation, and dephasing of individual spins at rates
-$\gamma_-=\gamma_+=\gamma_\z=100\chi$.  These results show that TNT is
-much more robust to this kind of decoherence than OAT or TAT, a
-finding that could not be deduced from the weak-decoherence
-simulations presented in Figure \ref{fig:benchmarking}.
-Strong-decoherence computations of this sort were used to put lower
-bounds on theoretically achievable spin squeezing via TAT with
-decoherence in Ref.~[\citenum{he2019engineering}], exemplifying a
-concrete and practical application of the TST expansion and the
-collective-spin structure constants calculated in this work.
+strong-decoherence regimes of large systems that are entirely
+inaccessible to other methods.  As an example, Figure
+\ref{fig:decoherence_strong} shows squeezing of $N=10^4$ spins
+initially in $\ket{\X}$, undergoing spontaneous decay, excitation, and
+dephasing of individual spins at rates
+$\gamma_-=\gamma_+=\gamma_\z=100\chi$.
+%
+% These results show that
+%
+The system size in these simulations is too large for straightforward
+application of exact methods for open quantum systems.  Quantum
+trajectory Monte Carlo simulations, meanwhile, take a prohibitively
+long time to converge with such strong decoherence due to the
+multiplicity of quantum trajectories that require averaging.
+
+The results in Figure \ref{fig:decoherence_strong} show that
+%
+TNT is much more robust to single-spin decoherence than OAT or TAT, a
+finding that could not necessarily be deduced from the
+weak-decoherence simulations presented in Figure
+\ref{fig:benchmarking}.  Strong-decoherence computations of this sort
+were used to put lower bounds on theoretically achievable spin
+squeezing via TAT with decoherence in
+Ref.~[\citenum{he2019engineering}], exemplifying a concrete and
+practical application of the TST expansion and the collective-spin
+structure constants calculated in this work.
 
@@
   D\p{t}
-  \equiv \f1{S^2}\Braket{\sp{\hat S_+\p{t},\hat S_-\p{0}}_-^\dag
-    \sp{\hat S_+\p{t},\hat S_-\p{0}}_-},
+  \equiv \f1{S^2} \Bk{\sp{\hat S_+\p{t},\hat S_-\p{0}}_-^\dag
+    \sp{\hat S_+\p{t},\hat S_-\p{0}}_-}_{\t{nn}},
   \label{eq:four_point}
@@
 in the context of the squeezing models in Section \ref{sec:squeezing}.
+%
+The subscript on $\bk{\cdot}_{\t{nn}}$ in Eq.~\eqref{eq:four_point}
+stands for ``no noise'', and denotes a correlator computed without the
+noise contributions $\hat\E_\O\p{t}$ to Heisenberg operators
+$\hat\O\p{t}$.  While linear contributions from noise operators as
+e.g.~in Eq.~\eqref{eq:two_time} generally vanish, quadratic
+contributions that would otherwise appear in Eq.~\eqref{eq:four_point}
+generally do not.  We exclude these contributions because their
+calculation generally requires additional assumptions about the
+environment (i.e.~beyond Markovianity) that would be a digression for
+the purposes of the present work; see Ref.~\cite{blocher2019quantum}
+for a detailed discussion of noise terms and the quantum regression
+theorem underlying the calculation of multi-time correlators.
+%
+
 In an equilibrium setting, correlation functions similar to that in
@@
 
-\begin{figure}
+\begin{figure*}
   \centering
+  \begin{minipage}{0.49\linewidth}
     \subfloat[Two-time correlator
-  $C\p{t}\equiv\abs{C\p{t}}\exp\sp{i\phi\p{t}}$. \label{fig:two_time}]
-  {\includegraphics{two_time.pdf}} \\
-  \subfloat[Squared commutator $D\p{t}$. \label{fig:four_point}]
+    $C\p{t}\equiv\abs{C\p{t}}\exp\sp{i\phi\p{t}}$.
+    \label{fig:two_time}]
+    {\includegraphics{two_time.pdf}}
+  \end{minipage}
+  \begin{minipage}{0.49\linewidth}
+    \subfloat[Squared commutator $D\p{t}$.
+    \label{fig:four_point}]
     {\includegraphics{FPC.pdf}}
+  \end{minipage}
   \caption{The two-time connected correlator $C\p{t}$ and squared
@@
   \label{fig:multi_time}
-\end{figure}
+\end{figure*}
 
@@
   We acknowledge helpful discussions with Robert Lewis-Swan, Kris
-  Tucker, Diego Barberena, and Colin Kennedy.  This work is supported
-  by the Air Force Office of Scientific Research (AFOSR) grant
-  FA9550-18-1-0319; the AFOSR Multidisciplinary University Research
-  Initiative (MURI) grant; the Defense Advanced Research Projects
-  Agency (DARPA) and Army Research Office (ARO) grant W911NF-16-1-0576
-  and W911NF-19-1-0210; the National Science Foundation (NSF) grant
-  PHY-1820885; JILA-NSF grant PFC-173400; and the National Institute
-  of Standards and Technology (NIST).
+  Tucker, and Colin Kennedy; as well as some technical contributions
+  from Diego Barberena.  This work is supported by the Air Force
+  Office of Scientific Research (AFOSR) grant FA9550-18-1-0319; the
+  AFOSR Multidisciplinary University Research Initiative (MURI) grant;
+  the Defense Advanced Research Projects Agency (DARPA) and Army
+  Research Office (ARO) grant W911NF-16-1-0576 and W911NF-19-1-0210;
+  the National Science Foundation (NSF) grant PHY-1820885; JILA-NSF
+  grant PFC-173400; and the National Institute of Standards and
+  Technology (NIST).
 \end{acknowledgments}
 
+\vfill
 \onecolumngrid
@@
 \begin{align}
-  \bk{S_+}
+  \bk{S_+\p{t}}
   &= S e^{-\Gamma t} \Phi\p{\chi,t}^{N-1}, \label{eq:S+_OAT} \\
-  \bk{S_\mu S_\z}
-  &= -\f{\mu}{2}\bk{S_\mu} + S \p{S-\f12} e^{-\Gamma t}
+  \bk{\p{S_\mu S_\z}\p{t}}
+  &= -\f{\mu}{2}\bk{S_\mu\p{t}} + S \p{S-\f12} e^{-\Gamma t}
   \Psi\p{\mu\chi,t} \Phi\p{\chi,t}^{N-2}, \\
-  \bk{S_\mu S_\nu}
-  &= \delta_{\mu,-\nu} \p{S + \mu\bk{S_\z}}
+  \bk{\p{S_\mu S_\nu}\p{t}}
+  &= \delta_{\mu,-\nu} \p{S + \mu\bk{S_\z\p{t}}}
   + S \p{S-\f12} e^{-2\Gamma t}
@@
 parameter $\xi^2$ defined in \eqref{eq:squeezing}, we additionally
-need analytical expressions for $\bk{S_\z}$ and $\bk{S_\z^2}$.  As
-these operators commute with both the OAT Hamiltonian and the
-single-spin operators $\sigma_\z^{(j)}$, their evolution is governed
-entirely by decay-type decoherence (see Appendix
+need analytical expressions for $\bk{S_\z\p{t}}$ and
+$\bk{S_\z^2\p{t}}$.  As these operators commute with both the OAT
+Hamiltonian and the single-spin operators $\sigma_\z^{(j)}$, their
+evolution is governed entirely by decay-type decoherence (see Appendix
 \ref{sec:decay_single}), which means
 \begin{align}
-  \f{d}{dt} S_\z
-  &= S\p{\gamma_+-\gamma_-} - \p{\gamma_++\gamma_-} S_\z,
+  \f{d}{dt} \bk{S_\z\p{t}}
+  &= S\p{\gamma_+-\gamma_-} - \p{\gamma_++\gamma_-} \bk{S_\z\p{t}},
   \\
-  \f{d}{dt}\p{S_\z^2}
-  &= S\p{\gamma_++\gamma_-} + 2\p{S-\f12}\p{\gamma_+-\gamma_-} S_\z
-  - 2 \p{\gamma_++\gamma_-} S_\z^2.
-\end{align}
-The initial conditions $\bk{S_\z}_{t=0}=0$ and $\bk{S_\z^2}_{t=0}=S/2$
+  \f{d}{dt} \bk{S_\z^2\p{t}}
+  &= S\p{\gamma_++\gamma_-}
+  + 2\p{S-\f12}\p{\gamma_+-\gamma_-} \bk{S_\z\p{t}}
+  - 2 \p{\gamma_++\gamma_-} \bk{S_\z^2\p{t}}.
+\end{align}
+The initial conditions $\bk{S_\z\p{0}}=0$ and $\bk{S_\z^2\p{0}}=S/2$
 then imply
 \begin{align}
-  \bk{S_\z} = S\p{\f{\gamma_+-\gamma_-}{\gamma_++\gamma_-}}
+  \bk{S_\z\p{t}} = S\p{\f{\gamma_+-\gamma_-}{\gamma_++\gamma_-}}
   \p{1-e^{-\p{\gamma_+ + \gamma_-} t}},
   &&
-  \bk{S_\z^2} = S \sp{\f12 + \p{S-\f12} \f{\bk{S_\z}^2}{S^2}}.
+  \bk{S_\z^2\p{t}}
+  = S \sp{\f12 + \p{S-\f12} \p{\f{\bk{S_\z\p{t}}}{S}}^2}.
   \label{eq:Sz_OAT}
@@
 Here we compute the expectation values of collective spin operators
-with respect to spin-polarized (Gaussian) states.  These states are
-parameterized by polar and azimuthal angles $\theta\in[0,\pi)$,
-$\phi\in[0,2\pi)$, and lie within the Dicke manifold spanned by states
+with respect to spin-polarized (also Gaussian, or spin-coherent)
+states.  These states are parameterized by polar and azimuthal angles
+$\theta\in[0,\pi)$, $\phi\in[0,2\pi)$, and lie within the Dicke
+manifold spanned by states
 $\ket{k}\propto S_+^{S+k}\ket{\dn}^{\otimes N}$ with $S\equiv N/2$ and
@@
 additionally subject to spontaneous single-spin decay at rate
-$\gamma_-$, with jump operators $\A_-=\set{s_-^{(j)}:j=1,2,\cdots,N}$.
+$\gamma_-$, with jump operators $\J_-=\set{s_-^{(j)}:j=1,2,\cdots,N}$.
 The equation of motion for a Heisenberg operator
@@
 \begin{align}
-  \f{d}{dt} \p{S_+^\ell S_\z^m S_-^n}
-  = i\chi\sp{S_\z^2, S_+^\ell S_\z^m S_-^n}_-
-  + \gamma_- \D\p{\A_-} \p{S_+^\ell S_\z^m S_-^n},
+  \f{d}{dt} \bk{S_+^\ell S_\z^m S_-^n}
+  = i\chi\Bk{\sp{S_\z^2, S_+^\ell S_\z^m S_-^n}_-}
+  + \gamma_- \Bk{\D\p{\J_-} \p{S_+^\ell S_\z^m S_-^n}},
   \label{eq:tutorial_EOM}
@@
 commutator $\sp{S_\z^2, S_+^\ell S_\z^m S_-^n}_-$ and dissipator
-$\D\p{\A_-} \p{S_+^\ell S_\z^m S_-^n}$ in \eqref{eq:tutorial_EOM} (see
+$\D\p{\J_-} \p{S_+^\ell S_\z^m S_-^n}$ in \eqref{eq:tutorial_EOM} (see
 \eqref{eq:general_product} and \eqref{eq:decay_diff} in particular),
 giving us
-\begin{align}
-  \f{d}{dt} \p{S_+^\ell S_\z^m S_-^n} =
-  i\chi \p{\ell-n} S_+^\ell \p{\ell+n+2S_\z} S_\z^m S_-^n
-  + \gamma_- S_+^\ell \sp{\p{S+S_\z}\p{-1+S_\z}^m
-    - \p{S+\f{\ell+n}{2}+S_\z} S_\z^m} S_-^n.
+\begin{multline}
+  \f{d}{dt} \bk{S_+^\ell S_\z^m S_-^n} \\
+  = i\chi \bk{\p{\ell-n} S_+^\ell \p{\ell+n+2S_\z} S_\z^m S_-^n}
+  + \gamma_- \Bk{S_+^\ell \sp{\p{S+S_\z}\p{-1+S_\z}^m
+      - \p{S+\f{\ell+n}{2}+S_\z} S_\z^m} S_-^n}.
   \label{eq:tutorial_EOM_expanded}
-\end{align}
+\end{multline}
 In practice, we do not want to keep track of such an expansion by
@@
 \begin{align}
-  \f{d}{dt} \S_{\v n} \equiv T \S_{\v n}
-  = \sum_{\v m} \S_{\v m} T_{\v m\v n},
+  \f{d}{dt} \bk{\S_{\v n}}
+  = \bk{T \S_{\v n}}
+  = \sum_{\v m} \bk{\S_{\v m}} T_{\v m\v n},
   \label{eq:tutorial_EOM_general}
@@
 For any Hamiltonian $H$ with decoherence characterized by sets of jump
-operators $\A$ and decoherence rates $\gamma_\A$, the matrix elements
+operators $\J$ and decoherence rates $\gamma_\J$, the matrix elements
 $T_{\v m\v n}$ are more generally defined by
 \begin{align}
-  T \S_{\v n}
-  = i\sp{H,\S_{\v n}}_- + \sum_\A \gamma_\A \D\p{\A} \S_{\v n}
-  = \sum_{\v m} \S_{\v m} T_{\v m\v n}.
+  \bk{T \S_{\v n}}
+  = i\bk{\sp{H,\S_{\v n}}_-}
+  + \sum_\J \gamma_\J \bk{\D\p{\J} \S_{\v n}}
+  = \sum_{\v m} \bk{\S_{\v m}} T_{\v m\v n}.
 \end{align}
@@
 $T_{\v m\v n}$, taking a particular Hamiltonian $H$ and decoherence
-processes $\set{\p{\A,\gamma_\A}}$ as inputs.
+processes $\set{\p{\J,\gamma_\J}}$ as inputs.
 
@@
   \bk{\S_{\v n}\p{t}}
+  = \sum_{k\ge0} \f{t^k}{k!} \bk{T^k\S_{\v n}\p{0}}
+  = \sum_{k\ge0} \f{t^k}{k!}
+  \sum_{\v m} \bk{\S_{\v m}\p{0}} T^k_{\v m\v n}
   \to \sum_{k=0}^M \f{t^k}{k!}
@@
 where $T^k_{\v m\v n}$ are matrix elements of the $k$-th time
-derivative operator, given by
+derivative operator $T^k$, given by
 \begin{align}
@@
   \equiv \f1{k!} \sum_{\v m} \bk{\S_{\v m}\p{0}} T^k_{\v m\v n},
+  \label{eq:tutorial_corr}
 \end{align}
@@
 Defining perturbed and unperturbed Heisenberg operators
-$\widetilde\O\p{t}\equiv e^{t\widetilde T}\O$ and
-$\O\p{t}\equiv e^{tT}\O$, we thus find that for sufficiently small
+$\widetilde\O\p{t}$ and $\O\p{t}$ that respectively satisfy
+$\bk{\widetilde\O\p{t}}=\bk{e^{t\widetilde T}\O}$ and
+$\bk{\O\p{t}}=\bk{e^{tT}\O}$, we thus find that for sufficiently small
 times $t$ and weak perturbations $V$,
 \begin{align}
-  \widetilde\O\p{t} - \O\p{t}
-  = \p{e^{t\widetilde T} - e^{tT}} \O
-  \approx \f12 i t \p{\sp{V,\O}_-\p{t} + \sp{V,\O\p{t}}_-}.
+  \bk{\widetilde\O\p{t} - \O\p{t}}
+  = \Bk{\p{e^{t\widetilde T} - e^{tT}} \O}
+  \approx \f12 i t \p{\bk{\sp{V,\O}_-\p{t}}
+    + \bk{\sp{V,\O\p{t}}_-}}.
   \label{eq:response}
 \end{align}
-Two-time operators $V\O\p{t}$ and $\O\p{t} V$, in addition to the
-Heisenberg operators $\p{V\O}\p{t}$ and $\p{\O V}\p{t}$, thus
-determine the short-time linear response of operators $\O\p{t}$ to
-perturbations $V$ of a Hamiltonian.
+Two-time correlators $\bk{V\O\p{t}}$ and $\bk{\O\p{t}V}$, in addition
+to the expectation values $\bk{\p{V\O}\p{t}}$ and
+$\bk{\p{\O V}\p{t}}$, thus determine the short-time linear response of
+correlators $\bk{\O\p{t}}$ to perturbations $V$ of a Hamiltonian.
 
@@
 %   \sum_{\substack{\p{\v j;\v m},\p{\v k;\v n} \\
-%       \abs{\set{j_\alpha^\mu}\cap\set{k_\beta^\nu}}=s}} X.
+%       \abs{\set{j_\a^\mu}\cap\set{k_\b^\nu}}=s}} X.
 %   \label{eq:SS_PP}
@@
 %   = m_\mu + n_\mu - \sum_\kappa \p{r_{\mu\kappa}+r_{\kappa\mu}}
-%   + \sum_{\alpha,\beta} \rho^{\alpha\beta}_\mu.
+%   + \sum_{\a,\b} \rho^{\a\b}_\mu.
 % \end{align}

### 2019-09-11 soc_squeezing/collective_paper/calculations_OAT.tex
added comments on cumulant expansion added more discussion of the breakdown of TST added appendix on Heisenberg operators in open quantum systems, and the 'noise' operators

--- 2019-05-21 soc_squeezing/collective_paper/collective_spin.tex
@@
-\documentclass[pra,twocolumn,longbibliography]{revtex4-2}
-
-% linking references
-\usepackage{hyperref}
-\hypersetup{
-  breaklinks=true,
-  colorlinks=true,
-  linkcolor=blue,
-  urlcolor=cyan,
-}
-
-%%% symbols, notations, etc.
-\usepackage{physics,braket,bm,commath,amssymb}
-\renewcommand{\t}{\text} % text in math mode
-\newcommand{\f}[2]{\dfrac{#1}{#2}} % shorthand for fractions
-\newcommand{\p}[1]{\left( #1 \right)} % parenthesis
-\renewcommand{\sp}[1]{\left[ #1 \right]} % square parenthesis
-\renewcommand{\v}{\bm} % bold vectors
-\newcommand{\uv}[1]{\bm{\hat{#1}}} % unit vectors
-
-\newcommand{\bk}{\braket} % shorthand for braket notation
-
-\newcommand{\A}{\mathcal{A}}
-\newcommand{\D}{\mathcal{D}}
-\newcommand{\F}{\mathcal{F}}
-\newcommand{\I}{\mathcal{I}}
-\renewcommand{\O}{\mathcal{O}}
-\renewcommand{\S}{\mathcal{S}}
-
-\newcommand{\C}{\mathbb{C}}
-\newcommand{\N}{\mathbb{N}}
-
-\newcommand{\z}{\text{z}}
-\newcommand{\x}{\text{x}}
-\newcommand{\y}{\text{y}}
-\newcommand{\Z}{\text{Z}}
-\newcommand{\X}{\text{X}}
-\newcommand{\Y}{\text{Y}}
-\newcommand{\bmu}{{\bar\mu}}
-\newcommand{\bnu}{{\bar\nu}}
-
-\usepackage{dsfont} % for identity operator
-\newcommand{\1}{\mathds{1}}
-
-\newcommand{\up}{\uparrow}
-\newcommand{\dn}{\downarrow}
-
-\renewcommand{\a}{\alpha} % free index
-\renewcommand{\b}{\beta} % free index
-
-%%% figures
-\usepackage{graphicx} % for figures
-\usepackage[caption=false]{subfig} % subfigures (via \subfloat[]{})
-\newcommand{\sref}[1]{\protect\subref{#1}} % for referencing subfigures
-\graphicspath{{./figures/}} % set path for all figures
-
-\usepackage[inline]{enumitem} % for inline enumeration
-
-\begin{document}
-
-\title{Short-time expansion of Heisenberg operators in open collective
-  quantum spin systems}
-
-\author{Michael A. Perlin}
-\email{mika.perlin@gmail.com}
-\author{Ana Maria Rey}
-\affiliation{JILA, National Institute of Standards and Technology and
-  University of Colorado, 440 UCB, Boulder, Colorado 80309, USA}
-\affiliation{Center for Theory of Quantum Matter, 440 UCB, Boulder,
-  Colorado 80309, USA}
-\affiliation{Department of Physics, University of Colorado, 390 UCB,
-  Boulder, Colorado 80309, USA}
-
-\begin{abstract}
-  We present a new method to compute short-time expectation values in
-  large collective spin systems with generic Markovian decoherence.
-  Our method is based on a Taylor expansion of a formal solution to
-  the equations of motion for Heisenberg operators.  This expansion
-  can be truncated at finite order to obtain virtually exact results
-  at short times that are relevant for metrological applications such
-  as spin squeezing.  In order to evaluate the expansion for
-  Heisenberg operators, we compute the relevant structure constants of
-  a collective spin operator algebra.  We demonstrate the utility of
-  our method by computing spin squeezing, two-time correlation
-  functions, and out-of-time-ordered correlators for $10^4$ spins in
-  strong-decoherence regimes that are otherwise inaccessible via
-  existing numerical methods.  Our method can be straightforwardly
-  generalized to the case of a collective spin coupled to bosonic
-  modes, relevant for trapped ion and cavity QED experiments, and may
-  be used to investigate short-time signatures of quantum chaos and
-  information scrambling.
-\end{abstract}
-
-\maketitle
-
-\section{Introduction}
-
-Collective spin systems are a versatile resource in quantum science
-for a range of applications including quantum-enhanced metrology and
-quantum simulation.  Theoretical interest in such systems dates back
-to the mid-twentieth century with the introduction of the Dicke
-model\cite{dicke1954coherence} that describes atoms cooperatively
-interacting with a single mode of a radiation field, and the
-Lipkin-Meshkov-Glick (LMG) model, a toy model for testing many-body
-approximation methods in contemporary nuclear
-physics\cite{lipkin1965validity, meshkov1965validity,
-  glick1965validity}.  On the experimental side, the development of
-advanced trapping, cooling, and control techniques in atomic,
-molecular, and optical (AMO) systems have enabled the realization of
-collective spin models in a broad range of platforms, including atomic
-ensembles\cite{takano2009spin, appel2009mesoscopic, chen2010heralded,
-  schleier-smith2010states, chen2011conditional}, Bose-Einstein
-condensates\cite{riedel2010atomchipbased, gross2010nonlinear,
-  klinder2015dynamical}, ultracold Fermi
-gasses\cite{martin2013quantum, smale2018observation}, trapped
-ions\cite{bohnet2016quantum}, optical
-cavities\cite{leroux2010implementation, bohnet2014reduced,
-  cox2016deterministic, hosten2016measurement, hosten2016quantum,
-  zhiqiang2017nonequilibrium, lewis-swan2018robust,
-  norcia2018cavitymediated}, and molecular
-sensors\cite{jones2009magnetic}.  These implementations have triggered
-studies of a variety of rich subjects, including quantum
-criticality\cite{latorre2005entanglement, alcalde2007functional,
-  morrison2008dissipationdriven, morrison2008collective,
-  sarandy2009classical, wang2012quantum, kessler2012dissipative,
-  majd2014lmg}, non-equilibrium
-phenomena\cite{walls1978nonequilibrium, morrison2008dynamical,
-  bhattacherjee2014nonequilibrium, klinder2015dynamical,
-  maghrebi2016nonequilibrium, zhiqiang2017nonequilibrium,
-  lang2018concurrence, smale2018observation}, and precision
-metrology\cite{wineland1992spin, itano1993quantum,
-  kitagawa1993squeezed, ma2011quantum, takano2009spin,
-  appel2009mesoscopic, chen2010heralded, schleier-smith2010states,
-  chen2011conditional, riedel2010atomchipbased, gross2010nonlinear,
-  zhong2010simplified, agarwal1997atomic, lau2014proposal,
-  huang2015quantum, wineland1992spin, kitagawa1993squeezed,
-  schleier-smith2010squeezing, leroux2010implementation,
-  bohnet2014reduced, cox2016deterministic, hosten2016measurement,
-  hosten2016quantum, lewis-swan2018robust, norcia2018cavitymediated,
-  jones2009magnetic, he2019engineering}.
-
-One of the primary motivations for studying collective spin systems is
-their application to quantum-enhanced metrology.  Quantum projection
-noise limits the error $\Delta\phi$ in the measurement of a phase
-angle $\phi$ with $N$ independent spins to
-$\Delta\phi\sim1/\sqrt{N}$\cite{wineland1992spin, itano1993quantum,
-  ma2011quantum}.  Collective spin systems enable a means to break
-through this limit via the preparation of many-body entangled states
-such as NOON\cite{jones2009magnetic, chen2010heralded,
-  zhong2010simplified}, spin-cat\cite{agarwal1997atomic,
-  lau2014proposal, huang2015quantum}, and spin-squeezed
-states\cite{wineland1992spin, kitagawa1993squeezed, takano2009spin,
-  appel2009mesoscopic, schleier-smith2010states,
-  schleier-smith2010squeezing, leroux2010implementation,
-  riedel2010atomchipbased, gross2010nonlinear, chen2011conditional,
-  bohnet2014reduced, cox2016deterministic, hosten2016measurement,
-  hosten2016quantum, lewis-swan2018robust, norcia2018cavitymediated,
-  ma2011quantum, he2019engineering} that allow for measurement errors
-$\Delta\phi\sim1/N^\varepsilon$ with $1/2<\varepsilon\le1$, where
-$\varepsilon=1$ saturates the Heisenberg
-limit\cite{zwierz2010general}.  Generating such states typically
-requires nonlinear dynamics that are realized using collisional,
-photon-mediated, or phonon-mediated interactions.  Although a truly
-collective spin model requires uniform, all-to-all interactions, as
-long as measurements do not distinguish between constituent particles,
-even non-uniform systems may be effectively described by a uniform
-model with renormalized parameters\cite{hu2015entangled}.
-
-In the absence of decoherence, permutation symmetry and total spin
-conservation divide the total Hilbert space of a collective spin
-system into superselection sectors that grow only linearly with system
-size $N$, thereby admitting efficient classical simulation of its
-dynamics.  Decoherence generally violates total spin conservation and
-requires the use of density operators, increasing the dimension of
-accessible state space to $\sim N^3$\cite{hartmann2016generalized,
-  xu2013simulating}.  In this case, exact simulations can be carried
-out for $N\lesssim100$ particles.  If decoherence is sufficiently
-weak, dynamics can be numerically solvable for $N\lesssim10^5$
-particles via ``quantum trajectory'' Monte Carlo
-methods\cite{plenio1998quantumjump, zhang2018montecarlo} (also known
-as ``quantum jump'' or ``Monte Carlo wavefunction'' methods) that can
-reproduce all expectation values of interest.  When decoherence is
-strong, however, or when the number of jump operators grows
-extensively with system size (e.g.~for single-spin decay), these Monte
-Carlo methods can take a prohibitively long time to converge, as
-simulations become dominated by incoherent jumps that generate large
-numbers of distinct quantum trajectories that need to be averaged in
-order to compute expectation values.
-
-In this work, we present a new method to compute short-time dynamics
-of collective spin systems with generic Markovian decoherence.  The
-only restriction on decoherence is that, like the coherent collective
-dynamics, it must act identically on all constituent particles.  This
-method is based on a short-time expansion of exact solutions to the
-equations of motion for Heisenberg operators.  Evaluating this
-expansion requires knowing the structure constants of a collective
-spin operator algebra; the calculation of these structure constants
-(in Appendices \ref{sec:identities}--\ref{sec:general_product}) is one
-of the main technical results of this work, which we hope will empower
-both analytical and numerical studies of collective spin systems in
-the future.  We benchmark our method against both exact and quantum
-trajectory Monte Carlo computations of spin squeezing in accessible
-parameter regimes, highlighting both advantages and limitations of our
-short-time expansion.  Finally, we showcase applications of our method
-with computations that are inaccessible to other numerical methods.
-
-\section{Theory}
-
-In this section we provide the basic theory for our method to compute
-expectation values of collective spin operators, deferring lengthy
-derivations to the appendices.  We consider systems composed of $N$
-distinct spin-1/2 particles.  Defining individual spin-1/2 operators
-$\hat s_{\a=\x,\y,\z}\equiv\hat\sigma_\a/2$ and
-$\hat s_\pm\equiv\hat s_\x\pm i\hat s_\y=\hat\sigma_\pm$ with Pauli
-operators $\hat\sigma_\a$, we denote an operator that acts with
-$\hat s_\a$ on the spin indexed by $j$ and trivially (i.e.~with the
-identity $\hat\1$) on all other spins by $\hat s_\a^{(j)}$.  We then
-define the collective spin operators
-$\hat S_\a\equiv\sum_{j=1}^N\hat s_\a^{(j)}$ for
-$\a\in\set{\x,\y,\z,+,-}$.  Identifying the set $\set{\hat\S_{\v m}}$
-as a basis for all collective spin operators, with
-$\v m\equiv\p{m_+,m_\z,m_-}\in\N_0^3$ and
-$\hat\S_{\v m}\equiv \hat S_+^{m_+} \hat S_\z^{m_\z} \hat S_-^{m_-}$,
-we can expand any collective spin operator $\hat\O$ in the form
-\begin{align}
-  \hat\O = \sum_{\v m} \O_{\v m} \hat\S_{\v m}
-\end{align}
-with scalar coefficients $\O_{\v m}\in\C$.  If $\hat\O$ is
-self-adjoint, for example, then $\O_{\v m}^*=\O_{\v m^*}$ with
-$\v m^*\equiv\p{m_-,m_\z,m_+}$.  The corresponding Heisenberg operator
-is then $\hat\O\p{t}=\sum_{\v m}\O_{\v m}\p{t}\hat\S_{\v m}$, with
-time-dependent coefficients $\O_{\v n}\p{t}$ for time-independent
-(Schr\"odinger) operators $\hat\S_{\v m}$.  The time evolution of
-Heisenberg operators is determined by
-\begin{align}
-  \f{d}{dt} \hat\O\p{t}
-  = \check T \hat\O\p{t}
-  = \sum_{\v m, \v n} \hat\S_{\v m} T_{\v m\v n} \O_{\v n}\p{t}
-  \label{eq:EOM}
-\end{align}
-for a time derivative (super-)operator $\check T$ with matrix elements
-$T_{\v m\v n}\in\C$ defined by
-\begin{align}
-  \check T \hat\S_{\v n} \equiv i \sp{\hat H, \hat\S_{\v n}}_-
-  + \sum_\A \gamma_\A \check\D\p{\A} \hat\S_{\v n}
-  = \sum_{\v m} \hat\S_{\v m} T_{\v m\v n},
-  \label{eq:time_deriv}
-\end{align}
-where $\sp{X,Y}_\pm\equiv XY\pm YX$; $\hat H$ is the collective spin
-Hamiltonian; $\A$ is a set of jump operators with a corresponding
-decoherence rate $\gamma_\A$; and $\check\D$ is a Heisenberg-picture
-dissipator, or Lindblad superoperator, defined by
-\begin{align}
-  \check\D\p{\A} \hat \O
-  \equiv \sum_{\hat A\in\A} \p{\hat A^\dag \hat \O \hat A
-    - \f12\sp{\hat A^\dag \hat A, \hat \O}_+}.
-\end{align}
-Decoherence via uncorrelated decay of individual spins, for example,
-would be described by the set of jump operators
-$\A_-\equiv\set{\hat s_-^{(j)}:j=1,2,\cdots,N}$.  The commutator in
-Eq.~\eqref{eq:time_deriv} can be computed by expanding the product
-$\hat\S_{\v\ell}\hat\S_{\v m} = \sum_{\v n} f_{\v\ell\v m\v
-  n}\hat\S_{\v n}$ with structure constants
-$f_{\v\ell\v m\v n}\in\mathbb{R}$ that we work out in Appendices
-\ref{sec:identities}--\ref{sec:general_product}, and the effects of
-decoherence from jump operators (i.e.~elements of $\A$) of the form
-$\hat g^{(j)} = \sum_\a g_\a \hat s_\a^{(j)}$ and
-$\hat G = \sum_\a G_\a \hat S_\a$ are worked out in Appendices
-\ref{sec:sandwich_single}--\ref{sec:decoherence_collective}.  We
-consider these calculations to be some of the main technical
-contributions of this work, with potential applications beyond the
-short-time simulation method presented here.  These ingredients are
-sufficient to compute matrix elements $T_{\v m\v n}$ of the time
-derivative operator $\check T$ in Eq.~\eqref{eq:time_deriv} in most
-cases of practical interest.
-
-The time derivative operator $\check T$ will generally couple spin
-operators $\hat\S_{\v n}$ to spin operators $\hat\S_{\v m}$ with
-higher ``weight'', i.e.~with $\abs{\v m}>\abs{\v n}$, where
-$\abs{\v\ell}\equiv\sum_\a\ell_\a$.  The growth of operator weight
-signifies the growth of many-body correlations.  In practice, keeping
-track of this growth will eventually require more computational
-resources than are available, meaning we must somehow truncate our
-equations of motion.  The simplest truncation strategy would be to
-take
-\begin{align}
-  \f{d}{dt} \hat\O\p{t}
-  \to \sum_{w\p{\v m}<W} \hat\S_{\v m}
-  \sum_{\v n} T_{\v m\v n} \O_{\v n}\p{t}
-  \label{eq:weight_truncation}
-\end{align}
-for some weight measure $w$, e.g.~$w\p{\v m}=\abs{\v m}$, and a
-high-weight cutoff $W$.  The truncation in
-Eq.~\eqref{eq:weight_truncation} closes the system of differential
-equations defined by Eq.~\eqref{eq:EOM}, and allows us to solve it
-using standard numerical methods.  Some initial conditions for this
-system of differential equations, namely expectation values of
-collective spin operators with respect to spin-polarized (Gaussian)
-states that are generally simple to prepare experimentally, are
-provided in Appendix \ref{sec:initial_conditions}.
-
-The truncation strategy in Eq.~\eqref{eq:weight_truncation} has a few
-limitations:
-\begin{enumerate*}[label=(\roman*)]
-\item simulating a system of differential equations for a large number
-  of operators can be time-consuming,
-\item the weight measure $w$ may need to be chosen carefully, as the
-  optimal measure is generally system-dependent, and
-\item simulation results can only be trusted up to the time at which
-  the initial value of operators $\hat\S_{\v m}$ with weight
-  $w\p{\v m}\ge W$ have a non-negligible contribution to expectation
-  values of interest.\label{pt:limitation}
-\end{enumerate*}
-The last limitation in particular unavoidably applies in some form to
-any method tracking only a subset of all relevant operators.  We
-therefore devise an alternate truncation strategy built around
-limitation \ref*{pt:limitation}.
-
-We can formally expand Heisenberg operators $\hat\O\p{t}$ in a Taylor
-series about the time $t=0$ to write
-\begin{align}
-  \hat\O\p{t}
-  = e^{t\check T} \hat \O\p{0}
-  = \sum_{k\ge0} \f{t^k}{k!}
-  \sum_{\v m, \v n} \hat\S_{\v m} T^k_{\v m\v n} \O_{\v n}\p{0},
-  \label{eq:time_series}
-\end{align}
-where the matrix elements $T^k_{\v m\v n}$ of the $k$-th time
-derivative operator $T^k$ are
-\begin{align}
-  T^0_{\v m\v n} &\equiv \delta_{\v m\v n}, \\
-  T^1_{\v m\v n} &\equiv T_{\v m\v n}, \\
-  T^{k>1}_{\v m\v n}
-  &\equiv \sum_{\v p_1,\v p_2,\cdots,\v p_{k-1}}
-  T_{\v m\v p_{k-1}} \cdots T_{\v p_3\v p_2}
-  T_{\v p_2\v p_1} T_{\v p_1\v n},
-\end{align}
-with $\delta_{\v m\v n}=1$ if $\v m=\v n$ and 0 otherwise.  For
-sufficiently short times, we can truncate the series in
-Eq.~\eqref{eq:time_series} by taking
-\begin{align}
-  \hat\O\p{t}
-  \to \sum_{k=0}^M \f{t^k}{k!}
-  \sum_{\v m, \v n} \hat\S_{\v m} T^k_{\v m\v n} \O_{\v n}\p{0}.
-  \label{eq:TST}
-\end{align}
-We refer Eq.~\eqref{eq:TST} as the truncated short-time (TST)
-expansion of Heisenberg operators.  Unlike the truncation in
-Eq.~\eqref{eq:weight_truncation}, the nonzero matrix elements
-$T^k_{\v m\v n}$ for $k=0,1,\cdots,M$ in Eq.~\eqref{eq:TST} tell us
-which operators are relevant for computing the expectation value
-$\bk{\hat\O\p{t}}$ to a fixed order $M$.  Note that using the relation
-$\hat\S_{\v m}^\dag=\hat\S_{\v m^*}$, which by Hermitian conjugation
-of Eq.~\eqref{eq:EOM} also implies that
-$T_{\v m^*\v n^*}=T_{\v m\v n}^*$, can cut both the number of
-initial-time expectation values $\bk{\hat\S_{\v m}}_{t=0}$ and the
-number of matrix elements $T_{\v m\v n}$ that we may need to
-explicitly compute roughly in half.
-
-In principle, the factorial suppression of terms at higher orders of
-the expansion in Eq.~\eqref{eq:time_series} implies for any given time
-$t$, there exists a truncation order $M_t$ for which the truncation
-error in Eq.~\eqref{eq:TST} is negligibly small.  In practice, only a
-maximal truncation order $M_{\t{max}}$ is accessible with limited
-computational resources, such that the TST expansion in
-Eq.~\eqref{eq:TST} only allows for computation of the expectation
-value $\bk{\hat\O\p{t}}$ to a maximal time $t_{M_{\t{max}}}$.  As we
-will see in the following section, $M_{\t{max}}=35$ will suffice for
-the computation of collective spin correlators up to times that are
-relevant e.g.~for spin squeezing protocols.  Appendix
-\ref{sec:tutorial} provides a pedagogical tutorial for computing
-correlators using the TST expansion.
-
-\section{Spin squeezing, benchmarking, and breakdown}
-\label{sec:squeezing}
-
-To benchmark our method for computing collective spin correlators, we
-consider three collective spin models known to generate spin-squeezed
-states: the one-axis twisting (OAT), two-axis twisting (TAT), and
-twist-and-turn (TNT) models described by the collective spin
-Hamiltonians\cite{kitagawa1993squeezed, micheli2003manyparticle}
-\begin{align}
-  H_{\t{OAT}} &= \chi \hat S_\z^2, \label{eq:OAT} \\
-  H_{\t{TAT}}
-  &= \f{\chi}{3} \p{\hat S_\z^2 - \hat S_\y^2}, \label{eq:TAT} \\
-  H_{\t{TNT}} &= \chi \hat S_\z^2 + \Omega \hat S_\x, \label{eq:TNT}
-\end{align}
-where we include a factor of $1/3$ in the TAT Hamiltonian because it
-naturally appears in realistic proposals to experimentally implement
-TAT\cite{liu2011spin, huang2015twoaxis}.  For simplicity, we further
-fix $\Omega=\chi S$ (with $S\equiv N/2$ throughout this work) to the
-critical value known to maximize the entanglement generation rate of
-TNT in the large-$N$ limit\cite{micheli2003manyparticle,
-  sorelli2015fast}.
-
-Note that the OAT model is a special case of the zero-field Ising
-model, whose quantum dynamics admits an exact analytic solution even
-in the presence of decoherence\cite{foss-feig2013nonequilibrium}.  We
-will benchmark our calculations using these analytical OAT results
-wherever applicable (see Appendix \ref{sec:OAT}, as well as the
-Supplementary Material of Ref.~[\citenum{bohnet2016quantum}]).
-
-The Hamiltonians in Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT} squeeze the
-initial product state $\ket{\X}\propto\p{\ket\up+\ket\dn}^{\otimes N}$
-with $\hat S_\x\ket{\X}=S\ket{\X}$.  Our measure of spin squeezing is
-the directionally-unbiased Ramsey squeezing parameter determined by
-the maximal gain in resolution $\Delta\phi$ of a phase angle $\phi$
-over that achieved by any spin-polarized product state
-(e.g.~$\ket{\X}$)\cite{wineland1992spin, ma2011quantum},
-\begin{align}
-  \xi^2
-  \equiv \f{\p{\Delta\phi_{\t{min}}}^2}{\p{\Delta\phi_{\t{polarized}}}^2}
-  = \f{N}{\lvert\braket{\uv S}\rvert^2}
-  \min_{\substack{\v v\perp\bk{\uv S}\\\v v\cdot\v v=1}}
-    \Braket{\p{\uv S\cdot\v v}^2},
-  \label{eq:squeezing}
-\end{align}
-where $\uv S\equiv\p{\hat S_\x,\hat S_\y,\hat S_\z}$ is a collective
-spin operator-valued vector, and the minimization is performed over
-all unit vectors $\v v$ orthogonal to the mean spin vector
-$\bk{\uv S}$.  This squeezing parameter is entirely determined by one-
-and two-spin correlators of the form $\bk{\hat S_\a}$ and
-$\bk{\hat S_\a \hat S_\b}$.  For the unitary dynamics discussed in
-this work, these correlators are obtainable via exact simulations of
-quantum dynamics in the $\p{N+1}$-dimensional Dicke manifold of states
-$\set{\ket{S,m}}$ with net spin $S$ and spin projection $m$ onto the
-$z$ axis, i.e.~with $\bk{S,m|\uv S^2|S,m}=S\p{S+1}$ and
-$\bk{S,m|\hat S_\z|S,m}=m$ for $m\in\set{-S,-S+1,\cdots,S}$.  In the
-presence of single-spin or collective decoherence, meanwhile, these
-correlators are obtainable with the collective-spin quantum trajectory
-Monte Carlo method developed in ref.~\cite{zhang2018montecarlo}.  In
-this work, these exact and quantum trajectory simulations will be used
-to benchmark the TST expansion in Eq.~\eqref{eq:TST}.
-
-\begin{figure}
-  \centering
-  \subfloat[Squeezing with unitary dynamics
-  \label{fig:squeezing_unitary}]
-  {\includegraphics{unitary.pdf}} \\
-  \subfloat[Squeezing with decoherence:
-  $\gamma_-=\gamma_+=\gamma_\z=\chi$
-  \label{fig:squeezing_incoherent}]
-  {\includegraphics{decoherence_weak.pdf}}
-  \caption{Spin squeezing of $N=10^4$ spins initially in $\ket\X$
-    under \sref{fig:squeezing_unitary} unitary and
-    \sref{fig:squeezing_incoherent} non-unitary dynamics, computed
-    using exact methods (solid lines), quantum trajectory simulations
-    (dots), and the TST expansion in Eq.~\eqref{eq:TST} with $M=35$
-    (dashed lines).  Solid circles mark the times at which the TST
-    expansion gives an unphysical result with $\xi^2<0$.}
-  \label{fig:benchmarking}
-\end{figure}
-
-Figure \ref{fig:benchmarking} compares the squeezing parameter $\xi^2$
-for $N=10^4$ spins initially in the state $\ket{\X}$ evolved under the
-Hamiltonians in Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT}, as computed via
-both benchmarking simulations and the TST expansion in
-Eq.~\eqref{eq:TST} with $M=35$.  Squeezing is shown for both unitary
-dynamics (Figure \ref{fig:squeezing_unitary}), as well as non-unitary
-dynamics in the presence of spontaneous decay, excitation, and
-dephasing of individual spins at rates $\chi$ (Figure
-\ref{fig:squeezing_incoherent}), respectively described by the sets of
-jump operators $\A_\a\equiv\set{\hat s_\a^{(j)}}$ with corresponding
-decoherence rates $\gamma_\a=\chi$ for $\a\in\set{-,+,\z}$.  The
-results shown in Figure \ref{fig:benchmarking} were computed in a
-rotated basis with
-$\p{\hat s_\z,\hat s_\x}\to\p{\hat s_\x,-\hat s_\z}$ and
-$\ket\X\to\ket{-\Z}\equiv\ket\dn^{\otimes N}$ (together with
-appropriate transformations of the Hamiltonian and jump operators);
-the only effect of this rotation on the results presented in Figure
-\ref{fig:benchmarking} is to prolong the time for which the TST
-expansion agrees with the benchmarking simulations.  The reason for
-different results in rotated basis has to do with the breakdown of the
-TST expansion, which we discuss further below.
-
-The main lesson from Figure \ref{fig:benchmarking} is that the TST
-expansion gives essentially exact results right up until a sudden and
-drastic departure that can be diagnosed by inspection.  The breakdown
-of the TST expansion in Figure \ref{fig:benchmarking} is marked by an
-unphysical squeezing parameter $\xi^2<0$, which occurs due to
-individual correlators taking unphysical values
-$\lvert\bk{\hat\S_{\v n}}\rvert\gtrsim S^{\abs{\v n}}$.  The sudden
-and drastic departure from virtually exact results occurs because the
-TST expansion neglects high-weight (i.e.~large-$\abs{\v m}$) operators
-$\hat\S_{\v m}$ whose contributions to a Heisenberg operator
-$\hat\O\p{t}$ of interest eventually become non-negligible.  For
-sufficiently large times $t$, these individual contributions may
-generally be large compared to the value of $\bk{\hat\O\p{t}}$, in
-which case their truncation will neglect terms that are large compared
-to the actual value of $\bk{\hat\O\p{t}}$.  This breakdown mechanism
-is also the reason for prolonged agreement between the TST expansion
-and benchmarking methods when squeezing the initial state $\ket{-\Z}$
-rather than $\ket{\X}$: for initial states $\ket{-\Z}$, all
-initial-time correlators $\bk{\hat\S_{\v m}}_{t=0}$ vanish unless
-$m_+=m_-=0$ (see Appendix \ref{sec:initial_conditions}), so there is a
-substantially smaller number of neglected terms with non-zero
-contribution to the correlator $\bk{\hat\O\p{t}}$.
-
-Although the TST expansion breaks down at short times, it has two key
-advantages over other methods to compute collective spin correlators.
-First, computing spin correlators with the TST expansion is generally
-much faster and requires much fewer computing resources than the
-alternatives.  The quantum trajectory Monte Carlo simulations
-performed for Figure \ref{fig:squeezing_incoherent}, for example, take
-$\sim10^4$ CPU hours to compute on standard modern computing hardware;
-the bulk of this time is spent performing sparse matrix-vector
-multiplication, leaving little room to further optimize run time.
-Parallelization can reduce the actual run time of these quantum
-trajectory simulations to $\sim10$ hours, but at the cost of greatly
-increasing computing resource requirements.  The TST expansion results
-in Figure \ref{fig:squeezing_incoherent}, meanwhile, take $\sim10$
-seconds to compute with a single CPU on similar hardware.
-
-\begin{figure}
-  \centering
-  \includegraphics{decoherence_strong.pdf}
-  \caption{Spin squeezing of $N=10^4$ spins initially in $\ket\X$ with
-    spontaneous decay, excitation, and dephasing of individual spins
-    at rates $\gamma_-=\gamma_+=\gamma_\z=100\chi$.  Computed using
-    the TST expansion in Eq.~\eqref{eq:TST} with $M=35$.  Solid
-    circles mark the times at which the TST expansion gives an
-    unphysical result with $\xi^2<0$.}
-  \label{fig:decoherence_strong}
-\end{figure}
-
-Second, the TST expansion enables computing spin correlators in
-strong-decoherence regimes entirely inaccessible to other methods.  As
-an example, Figure \ref{fig:decoherence_strong} shows squeezing of
-$N=10^4$ spins initially in $\ket{\X}$, undergoing spontaneous decay,
-excitation, and dephasing of individual spins at rates
-$\gamma_-=\gamma_+=\gamma_\z=100\chi$.  These results show that TNT is
-much more robust to this kind of decoherence than OAT or TAT, a
-finding that could not be deduced from the weak-decoherence
-simulations presented in Figure \ref{fig:benchmarking}.
-Strong-decoherence computations of this sort were used to put lower
-bounds on theoretically achievable spin squeezing via TAT with
-decoherence in Ref.~[\citenum{he2019engineering}], exemplifying a
-concrete and practical application of the TST expansion and the
-collective-spin structure constants calculated in this work.
-
-\section{Two-time correlation functions and out-of-time-ordered
-  correlators}
-\label{sec:multi_time}
-
-As a final example of collective-spin physics that is numerically
-accessible via the TST expansion of Heisenberg operators, we consider
-the calculation of two-time correlation functions and
-out-of-time-ordered correlators (OTOCs).  In particular, we consider
-the effect of decoherence on short-time behavior of the two-time
-connected correlator
-\begin{align}
-  C\p{t}
-  \equiv \f1S\p{\bk{\hat S_+\p{t} \hat S_-\p{0}}
-    - \bk{\hat S_+\p{t}}\bk{\hat S_-\p{0}}},
-  \label{eq:two_time}
-\end{align}
-and the expectation value of a squared commutator,
-\begin{align}
-  D\p{t}
-  \equiv \f1{S^2}\Braket{\sp{\hat S_+\p{t},\hat S_-\p{0}}_-^\dag
-    \sp{\hat S_+\p{t},\hat S_-\p{0}}_-},
-  \label{eq:four_point}
-\end{align}
-in the context of the squeezing models in Section \ref{sec:squeezing}.
-In an equilibrium setting, correlation functions similar to that in
-Eq.~\eqref{eq:two_time} contain information about the linear response
-of Heisenberg operators to perturbations of a system; in a
-non-equilibrium setting, they contribute to short-time linear response
-(see Appendix \ref{sec:linear_response}).  Similar correlators have
-made appearances as order parameters for diagnosing time-crystalline
-phases of matter\cite{tucker2018shattered}.  Squared commutators such
-as that in Eq.~\eqref{eq:four_point}, meanwhile, are commonly examined
-for signatures of quantum chaos and information
-scrambling\cite{maldacena2016bound, swingle2018unscrambling,
-  garcia-mata2018chaos}.  In typical scenarios, such squared
-commutators initially vanish by construction due to a choice of
-operators with spatially separated non-trivial support.  Collective
-spin systems, however, have no intrinsic notion of locality or spatial
-separation.  In our case, therefore, with the choice of initial state
-$\ket\X\propto\p{\ket\up+\ket\dn}^{\otimes N}$ we merely have
-$D\p{0}\sim1/N$.
-
-\begin{figure}
-  \centering
-  \subfloat[Two-time correlator
-  $C\p{t}\equiv\abs{C\p{t}}\exp\sp{i\phi\p{t}}$. \label{fig:two_time}]
-  {\includegraphics{two_time.pdf}} \\
-  \subfloat[Squared commutator $D\p{t}$. \label{fig:four_point}]
-  {\includegraphics{FPC.pdf}}
-  \caption{The two-time connected correlator $C\p{t}$ and squared
-    commutator $D\p{t}$, respectively defined in
-    Eqs.~\eqref{eq:two_time} and \eqref{eq:four_point}, for $N=10^4$
-    spins initially in the polarized state
-    $\ket{\X}\propto\p{\ket\up+\ket\dn}^{\otimes N}$ evolving under
-    the squeezing Hamiltonians in Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT}.
-    Results are shown for both unitary dynamics (solid lines) and
-    non-unitary dynamics with $\gamma_-=\gamma_+=\gamma_\z=100\chi$
-    (dashed lines), computed using the TST expansion in
-    Eq.~\eqref{eq:TST} with $M=20$.}
-  \label{fig:multi_time}
-\end{figure}
-
-Figure \ref{fig:multi_time} shows the behavior of $C\p{t}$ and
-$D\p{t}$ for $N=10^4$ spins, initially in the state $\ket\X$, evolving
-under the squeezing Hamiltonians in
-Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT} both with and without spontaneous
-decay, excitation, and dephasing of individual spins at rates
-$\gamma_-=\gamma_+=\gamma_\z=100\chi$.  In the case of unitary
-evolution under OAT, we find that to an excellent approximation
-$\abs{C\p{t}}$ takes the functional form
-$f\p{t}=f\p{0}+aN\chi t+\p{bN\chi t}^2$ with $a\sim b\sim 1$, and with
-a virtually perfect fit $D\p{t}=D\p{0}+\p{\sp{N+1}\chi t}^2$.  For
-unitary evolution under TAT and TNT, we find that to an excellent
-approximation both $\abs{C\p{t}}$ and $D\p{t}$ take the functional
-form $f\p{t}=f\p{0}+a\sp{\exp\p{bN\chi t}-1}$ with $a\sim b\sim 1$.
-As may be expected, the growth of $C\p{t}$ and $D\p{t}$ is generally
-suppressed by decoherence.  Figure \ref{fig:multi_time} serves as an
-example for the type of behavior that is accessible at short times
-with the TST expansion.  These examples are straightforward extend to
-equilibrium settings and spin-boson systems.
-
-\section{Conclusions}
-
-We have presented a new method for computing correlators at short
-times in collective spin systems.  This method is based on truncating
-a short-time expansion of Heisenberg operators, and can access
-correlators on time scales that are relevant to metrological
-applications such as spin squeezing.  In order to evaluate the
-truncated short-time (TST) expansion of Heisenberg operators, we have
-computed the structure constants of a collective spin operator
-algebra, which we hope will empower future analytical and numerical
-studies of collective spin systems.  Even though we considered only
-non-equilibrium spin-squeezing processes in this work, our method can
-be applied directly in an equilibrium setting, and is straightforward
-to generalize to systems such as trapped ions and optical cavities
-with collective spin-boson interactions.  In such contexts, our method
-may be used to benchmark the short-time effects of decoherence, or
-study the onset of quantum chaos and information scrambling.
-
-\begin{acknowledgments}
-  We acknowledge helpful discussions with Robert Lewis-Swan, Kris
-  Tucker, Diego Barberena, and Colin Kennedy.  This work is supported
-  by the Air Force Office of Scientific Research (AFOSR) grant
-  FA9550-18-1-0319; the AFOSR Multidisciplinary University Research
-  Initiative (MURI) grant; the Defense Advanced Research Projects
-  Agency (DARPA) and Army Research Office (ARO) grant W911NF-16-1-0576
-  and W911NF-19-1-0210; the National Science Foundation (NSF) grant
-  PHY-1820885; JILA-NSF grant PFC-173400; and the National Institute
-  of Standards and Technology (NIST).
-\end{acknowledgments}
-
-\onecolumngrid
-\appendix
-
 \section{Analytical results for the one-axis twisting model}
@@
 \begin{align}
-  \bk{S_+}
+  \bk{S_+\p{t}}
   &= S e^{-\Gamma t} \Phi\p{\chi,t}^{N-1}, \label{eq:S+_OAT} \\
-  \bk{S_\mu S_\z}
-  &= -\f{\mu}{2}\bk{S_\mu} + S \p{S-\f12} e^{-\Gamma t}
+  \bk{\p{S_\mu S_\z}\p{t}}
+  &= -\f{\mu}{2}\bk{S_\mu\p{t}} + S \p{S-\f12} e^{-\Gamma t}
   \Psi\p{\mu\chi,t} \Phi\p{\chi,t}^{N-2}, \\
-  \bk{S_\mu S_\nu}
-  &= \delta_{\mu,-\nu} \p{S + \mu\bk{S_\z}}
+  \bk{\p{S_\mu S_\nu}\p{t}}
+  &= \p{1-\delta_{\mu\nu}} \p{S + \mu\bk{S_\z\p{t}}}
   + S \p{S-\f12} e^{-2\Gamma t}
@@
 parameter $\xi^2$ defined in \eqref{eq:squeezing}, we additionally
-need analytical expressions for $\bk{S_\z}$ and $\bk{S_\z^2}$.  As
-these operators commute with both the OAT Hamiltonian and the
-single-spin operators $\sigma_\z^{(j)}$, their evolution is governed
-entirely by decay-type decoherence (see Appendix
+need analytical expressions for $\bk{S_\z\p{t}}$ and
+$\bk{S_\z^2\p{t}}$.  As these operators commute with both the OAT
+Hamiltonian and the single-spin operators $\sigma_\z^{(j)}$, their
+evolution is governed entirely by decay-type decoherence (see Appendix
 \ref{sec:decay_single}), which means
 \begin{align}
-  \f{d}{dt} S_\z
-  &= S\p{\gamma_+-\gamma_-} - \p{\gamma_++\gamma_-} S_\z,
+  \f{d}{dt} \bk{S_\z\p{t}}
+  &= S\p{\gamma_+-\gamma_-} - \p{\gamma_++\gamma_-} \bk{S_\z\p{t}},
   \\
-  \f{d}{dt}\p{S_\z^2}
-  &= S\p{\gamma_++\gamma_-} + 2\p{S-\f12}\p{\gamma_+-\gamma_-} S_\z
-  - 2 \p{\gamma_++\gamma_-} S_\z^2.
+  \f{d}{dt} \bk{S_\z^2\p{t}}
+  &= S\p{\gamma_++\gamma_-}
+  + 2\p{S-\f12}\p{\gamma_+-\gamma_-} \bk{S_\z\p{t}}
+  - 2 \p{\gamma_++\gamma_-} \bk{S_\z^2\p{t}}.
 \end{align}
-The initial conditions $\bk{S_\z}_{t=0}=0$ and $\bk{S_\z^2}_{t=0}=S/2$
+The initial conditions $\bk{S_\z\p{0}}=0$ and $\bk{S_\z^2\p{0}}=S/2$
 then imply
 \begin{align}
-  \bk{S_\z} = S\p{\f{\gamma_+-\gamma_-}{\gamma_++\gamma_-}}
+  \bk{S_\z\p{t}} = S\p{\f{\gamma_+-\gamma_-}{\gamma_++\gamma_-}}
   \p{1-e^{-\p{\gamma_+ + \gamma_-} t}},
   &&
-  \bk{S_\z^2} = S \sp{\f12 + \p{S-\f12} \f{\bk{S_\z}^2}{S^2}}.
+  \bk{S_\z^2\p{t}}
+  = S \sp{\f12 + \p{S-\f12} \p{\f{\bk{S_\z\p{t}}}{S}}^2}.
   \label{eq:Sz_OAT}
@@
 state $\ket\X$ under $H_{\t{OAT}}$.
-
-\section{Basic spin operator identities}
-\label{sec:identities}
-
-The appendices in this work make ubiquitous use of various spin
-operator identities; we collect and derive some basic identities here
-for reference.  Note that despite the working definition of collective
-spin operators from $S_\a=\sum_js_\a^{(j)}$, the identities we will
-derive involving only collective spin operators apply just as well to
-large-spin operators that cannot be expressed as the sum of individual
-spin-1/2 operators.  The elementary commutation relations between spin
-operators are, with $\bmu\equiv-\mu\in\set{+1,-1}$ for brevity,
-\begin{align}
-  \sp{s_\z^{(j)},s_\mu^{(k)}}_-
-  &= \delta_{jk} \mu s_\mu^{(j)},
-  &
-  \sp{S_\z,s_\mu^{(j)}}_-
-  &= \sp{s_\z^{(j)},S_\mu}_- = \mu s_\mu^{(j)},
-  &
-  \sp{S_\z,S_\mu}_-
-  &= \mu S_\mu,
-  \label{eq:comm_z_base} \\
-  \sp{s_\mu^{(j)},s_\bmu^{(k)}}_-
-  &= \delta_{jk} 2 \mu s_\z^{(j)},
-  &
-  \sp{S_\mu,s_\bmu^{(j)}}_-
-  &= \sp{s_\mu^{(j)},S_\bmu}_- = 2 \mu s_\z^{(j)},
-  &
-  \sp{S_\mu,S_\bmu}_-
-  &= 2 \mu S_\z.
-  \label{eq:comm_mu_base}
-\end{align}
-These relations can be used to inductively compute identities
-involving powers of collective spin operators.  By pushing through one
-spin operator at a time, we can find
-\begin{align}
-  \p{\mu S_\z}^m s_\mu^{(j)}
-  = \p{\mu S_\z}^{m-1} s_\mu^{(j)} \p{1 + \mu S_\z}
-  = \p{\mu S_\z}^{m-2} s_\mu^{(j)} \p{1 + \mu S_\z}^2
-  = \cdots
-  = s_\mu^{(j)} \p{1 + \mu S_\z}^m,
-  \label{eq:push_z_mu_Ss}
-\end{align}
-and
-\begin{align}
-  \mu s_\z^{(j)} S_\mu^m
-  = S_\mu \mu s_\z^{(j)} S_\mu^{m-1} + s_\mu^{(j)} S_\mu^{m-1}
-  = \cdots
-  = S_\mu^m \mu s_\z^{(j)} + ms_\mu^{(j)} S_\mu^{m-1},
-  \label{eq:push_z_mu_sS}
-\end{align}
-where we will generally find it nicer to express results in terms of
-$\mu s_\z^{(j)}$ and $\mu S_\z$ rather than $s_\z^{(j)}$ and $S_\z$.
-Summing over the single-spin index $j$ in both of the cases above
-gives us the purely collective-spin versions of these identities:
-\begin{align}
-  \p{\mu S_\z}^m S_\mu = S_\mu \p{1 + \mu S_\z}^m,
-  &&
-  \mu S_\z S_\mu^m = S_\mu^m \p{m + \mu S_\z},
-  \label{eq:push_z_mu_single}
-\end{align}
-where we can repeat the process of pushing through individual $S_\z$
-operators $\ell$ times to get
-\begin{align}
-  \p{\mu S_\z}^\ell S_\mu^m
-  = \p{\mu S_\z}^{\ell-1} S_\mu^m \p{m + \mu S_\z}
-  = \p{\mu S_\z}^{\ell-2} S_\mu^m \p{m + \mu S_\z}^2
-  = \cdots
-  = S_\mu^m \p{m + \mu S_\z}^\ell.
-  \label{eq:push_z_mu}
-\end{align}
-Multiplying \eqref{eq:push_z_mu} through by $\p{\mu\nu}^\ell$ (for
-$\nu\in\set{+1,-1}$) and taking its Hermitian conjugate, we can say
-that more generally
-\begin{align}
-  \p{\nu S_\z}^\ell S_\mu^m
-  = S_\mu^m \p{\mu\nu m+\nu S_\z}^\ell,
-  &&
-  S_\mu^m \p{\nu S_\z}^\ell
-  = \p{-\mu\nu m+\nu S_\z}^\ell S_\mu^m.
-\end{align}
-Finding commutation relations between powers of transverse spin
-operators, i.e.~$S_\mu$ and $S_\bmu$, turns out to be considerably
-more difficult than the cases we have worked out thus far.  We
-therefore save this work for Appendix \ref{sec:comm_transverse}.
-
-\section{Commutation relations between powers of transverse spin
-  operators}
-\label{sec:comm_transverse}
-
-To find commutation relations between powers of transverse collective
-spin operators, we first compute
-\begin{align}
-  S_\mu^m s_\bmu^{(j)}
-  &= S_\mu^{m-1}s_\bmu^{(j)} S_\mu
-  + S_\mu^{m-1} 2\mu s_\z^{(j)} \\
-  &= S_\mu^{m-2} s_\bmu^{(j)} S_\mu^2
-  + S_\mu^{m-2} 2\mu s_\z^{(j)} S_\mu
-  + S_\mu^{m-1} 2\mu s_\z^{(j)} \\
-  &= s_\bmu^{(j)} S_\mu^m
-  + \sum_{k=0}^{m-1} S_\mu^k 2\mu s_\z^{(j)} S_\mu^{m-k-1}
-  \label{eq:push_mu_Ss_start}.
-\end{align}
-While \eqref{eq:push_mu_Ss_start} gives us the commutator
-$\sp{S_\mu^m,s_\bmu^{(j)}}_-$, we would like to enforce an ordering on
-products of spin operators, which will ensure that we only keep track
-of operators that are linearly independent.  We choose (for now) to
-impose an ordering with all $s_\bmu^{(j)}$ operators on the left, and
-all $s_\z^{(j)}$ operators on the right.  Such an ordering will prove
-convenient for the calculations in this section\footnote{In
-  retrospect, it may have been nicer to push all $s_\mu^{(j)}$
-  operators to the right throughout these calculations, due to the
-  enhanced symmetry expressions would have with respect to Hermitian
-  conjugation.  In any case, we provide the final result of this
-  section in both ordering conventions, and therefore feel no need to
-  reproduce these calculations with a different ordering of spin
-  operators.}.  This choice of ordering compels us to expand
-\begin{align}
-  \sum_{k=0}^{m-1} S_\mu^k 2\mu s_\z^{(j)} S_\mu^{m-k-1}
-  &= \sum_{k=0}^{m-1} S_\mu^k
-  \sp{2\p{m-k-1} s_\mu^{(j)} S_\mu^{m-k-2}
-    + S_\mu^{m-k-1} 2\mu s_\z^{(j)}} \\
-  &= m \p{m-1} s_\mu^{(j)} S_\mu^{m-2}
-  + m S_\mu^{m-1} 2\mu s_\z^{(j)},
-\end{align}
-which implies
-\begin{align}
-  S_\mu^m s_\bmu^{(j)}
-  = s_\bmu^{(j)} S_\mu^m + m \p{m-1} s_\mu^{(j)} S_\mu^{m-2}
-  + m S_\mu^{m-1} 2\mu s_\z^{(j)},
-  \label{eq:push_mu_Ss}
-\end{align}
-and in turn
-\begin{align}
-  S_\mu^m S_\bmu = S_\bmu S_\mu^m
-  + m S_\mu^{m-1} \p{m - 1 + 2\mu S_\z}.
-  \label{eq:push_mu_single}
-\end{align}
-As the next logical step, we take on the task of computing
-\begin{align}
-  S_\mu^m S_\bmu^n
-  = S_\mu^{m-1} S_\bmu^n S_\mu
-  + n \sp{S_\mu^{m-1} S_\bmu^{n-1} \p{1 - n + 2\mu S_\z}}
-  = S_\bmu^n S_\mu^m
-  + n \sum_{k=0}^{m-1} S_\mu^{m-k-1} S_\bmu^{n-1}
-  \p{1 - n + 2\mu S_\z} S_\mu^k,
-\end{align}
-which implies
-\begin{align}
-  \sp{S_\mu^m, S_\bmu^n}_-
-  = C_{mn;\mu}
-  \equiv n \sum_{k=0}^{m-1} S_\mu^{m-k-1} S_\bmu^{n-1}
-  \p{1 - n + 2\mu S_\z} S_\mu^k.
-\end{align}
-We now need rearrange the operators in $C_{mn;\mu}$ into a standard
-order, which means pushing all $S_\z$ operators to the right and, for
-the purposes of this calculation, all $S_\bmu$ operators to the left.
-We begin by pushing $S_\mu^k$ to the left of $S_\z$, which takes
-$2\mu S_\z\to 2\mu S_\z+2k$, and then push $S_\mu^{m-k-1}$ to the
-right of $S_\bmu^{n-1}$, giving us
-\begin{align}
-  C_{mn;\mu}
-  &= n \sum_{k=0}^{m-1}
-  \p{S_\bmu^{n-1} S_\mu^{m-k-1} + C_{m-k-1,n-1;\mu}} S_\mu^k
-  \p{2k + 1 - n + 2\mu S_\z} \\
-  &= D_{mn;\mu}
-  + n \sum_{k=0}^{m-2} C_{m-k-1,n-1;\mu}
-  S_\mu^k \p{2k + 1 - n + 2\mu S_\z},
-  \label{eq:C_mn}
-\end{align}
-where we have dropped the last ($k=m-1$) term in the remaining sum
-because $C_{m-k-1,n-1;\mu}=0$ if $k=m-1$, and
-\begin{align}
-  D_{mn;\mu}
-  \equiv mn S_\bmu^{n-1} S_\mu^{m-1} \p{m - n + 2\mu S_\z}.
-  \label{eq:D_mn}
-\end{align}
-To our despair, we have arrived in \eqref{eq:C_mn} at a {\it
-  recursive} formula for $C_{mn;\mu}$.  Furthermore, we have not even
-managed to order all spin operators, as $C_{m-k-1,n-1;\mu}$ contains
-$S_\z$ operators that are to the left of $S_\mu^k$.  To sort all spin
-operators once and for all, we define
-\begin{align}
-  C_{mn;\mu}^{(k)} \equiv C_{m-k,n;\mu} S_\mu^k,
-  &&
-  D_{mn;\mu}^{(k)} \equiv D_{m-k,n;\mu} S_\mu^k,
-\end{align}
-which we can expand as
-\begin{align}
-  D_{mn;\mu}^{(k)}
-  &= \p{m-k}n S_\bmu^{n-1} S_\mu^{m-k-1}
-  \p{m-k-n+2\mu S_\z} S_\mu^k \\
-  &= \p{m-k}n S_\bmu^{n-1} S_\mu^{m-1} \p{k+m-n+2\mu S_\z},
-  \label{eq:D_mn_k}
-\end{align}
-and
-\begin{align}
-  C_{mn;\mu}^{(k)}
-  &= D_{m-k,n;\mu} S_\mu^k + n \sum_{j=0}^{m-k-2}
-  C_{m-k-j-1,n-1;\mu} S_\mu^j \p{2j+1-n+2\mu S_\z} S_\mu^k \\
-  &= D_{mn;\mu}^{(k)} + n \sum_{j=0}^{m-k-2}
-  C_{m-k-j-1,n-1;\mu} S_\mu^{j+k} \p{2j+2k+1-n+2\mu S_\z} \\
-  &= D_{mn;\mu}^{(k)} + n \sum_{j=0}^{m-k-2}
-  C_{m-1,n-1;\mu}^{(k+j)} \p{2\sp{j+k}+1-n+2\mu S_\z} \\
-  &= D_{mn;\mu}^{(k)} + n \sum_{j=k}^{m-2}
-  C_{m-1,n-1;\mu}^{(j)} \p{2j+1-n+2\mu S_\z}.
-  \label{eq:C_mn_k}
-\end{align}
-While the resulting expression in \eqref{eq:C_mn_k} strongly resembles
-that in \eqref{eq:C_mn}, there is one crucial difference: all spin
-operators in \eqref{eq:C_mn_k} have been sorted into a standard order.
-We can now repeatedly substitute $C_{mn;\mu}^{(k)}$ into itself, each
-time decreasing $m$ and $n$ by 1, until one of $m$ or $n$ reaches
-zero.  Such repeated substitution yields the expansion
-\begin{align}
-  C_{mn;\mu}
-  = C_{mn;\mu}^{(0)}
-  = D_{mn;\mu}
-  + \sum_{p=1}^{\min\set{m,n}-1} E_{mn;\mu}^{(p)},
-  \label{eq:C_mn_E}
-\end{align}
-where the first two terms in the sum over $p$ are
-\begin{align}
-  E_{mn;\mu}^{(1)}
-  &= n \sum_{k=0}^{m-2} D_{m-1,n-1;\mu}^{(k)} \p{2k+1-n+2\mu S_\z}, \\
-  E_{mn;\mu}^{(2)}
-  &= n \sum_{k_1=0}^{m-2} \p{n-1} \sum_{k_2=k_1}^{m-3}
-  D_{m-2,n-2;\mu}^{(k_2)} \p{2k_2+2-n+2\mu S_\z} \p{2k_1+1-n+2\mu S_\z},
-\end{align}
-and more generally for $p>1$,
-\begin{align}
-  E_{mn;\mu}^{(p)}
-  = \f{n!}{\p{n-p}!}
-  \sum_{k_1=0}^{m-2} \sum_{k_2=k_1}^{m-3} \cdots\sum_{k_p=k_{p-1}}^{m-p-1}
-  D_{m-p,n-p;\mu}^{(k_p)} \prod_{j=1}^p \p{2k_j+j-n+2\mu S_\z}.
-  \label{eq:E_mn_p}
-\end{align}
-In principle, the expressions in \eqref{eq:D_mn}, \eqref{eq:D_mn_k},
-\eqref{eq:C_mn_E}, and \eqref{eq:E_mn_p} suffice to evaluate the
-commutator $\sp{S_\mu^m,S_\bmu^n}_- = C_{mn;\mu}$, but this result is
--- put lightly -- quite a mess: the expression for $E_{mn;\mu}^{(p)}$
-in \eqref{eq:E_mn_p} involves a sum over $p$ mutually dependent
-intermediate variables, each term of which additionally contains a
-product of $p$ factors.  We therefore devote the rest of this section
-to simplifying our result for the commutator
-$\sp{S_\mu^m,S_\bmu^n}_-$.
-
-Observing that in \eqref{eq:E_mn_p} we always have
-$0\le k_1\le k_2\le\cdots\le k_p\le m-p-1$, we can rearrange the order
-of the sums and relabel $k_p\to\ell$ to get
-\begin{align}
-  E_{mn;\mu}^{(p)}
-  = \f{n!}{\p{n-p}!}
-  \sum_{\ell=0}^{m-p-1} D_{m-p,n-p;\mu}^{(\ell)} \p{2\ell+F_{np;\mu}}
-  \sum_{\p{\v k,p-1,\ell}} \prod_{j=1}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}},
-  \label{eq:E_mn_p_sum}
-\end{align}
-where for shorthand we define
-\begin{align}
-  F_{np;\mu} \equiv p - n + 2\mu S_\z,
-  &&
-  \sum_{\p{\v k,q,\ell}} X \equiv
-  \sum_{k_1=0}^\ell \sum_{k_2=k_1}^\ell
-  \cdots \sum_{k_q=k_{q-1}}^\ell X.
-\end{align}
-We now further define
-\begin{align}
-  f_{np\ell;\mu}\p{k,q} \equiv \p{\ell-k+q} \p{\ell+k-q+F_{np;\mu}},
-\end{align}
-and evaluate sums successively over $k_{p-1},k_{p-2},\cdots,k_1$,
-finding
-\begin{align}
-  \sum_{\p{\v k,p-1,\ell}} \prod_{j=1}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}}
-  &= \sum_{\p{\v k,p-2,\ell}}
-  \prod_{j=2}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}}
-  f_{np\ell;\mu}\p{k_{p-2},1} \\
-  &= \f1{\p{r-1}!} \sum_{\p{\v k,p-r,\ell}}
-  \prod_{j=r}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}}
-  \prod_{q=1}^{r-1} f_{np\ell;\mu}\p{k_{p-r},q} \\
-  &= \f1{\p{p-1}!} \prod_{q=1}^{p-1} f_{np\ell;\mu}\p{0,q} \\
-  &= { \ell + p - 1 \choose p - 1 }
-  \prod_{q=1}^{p-1} \p{\ell-q+F_{np;\mu}}.
-\end{align}
-Substitution of this result together with $D_{m-p,n-p;\mu}^{(\ell)}$
-using \eqref{eq:D_mn_k} into \eqref{eq:E_mn_p_sum} then gives us
-\begin{align}
-  E_{mn;\mu}^{(p)}
-  = \f{n!}{\p{n-p-1}!} S_\bmu^{n-p-1} S_\mu^{m-p-1} G_{mnp;\mu}
-\end{align}
-with
-\begin{align}
-  G_{mnp;\mu}
-  &\equiv \sum_{\ell=0}^{m-p-1} { \ell + p - 1 \choose p - 1 }
-  \p{m-p-\ell} \p{\ell+m-p+F_{np;\mu}}
-  \p{2\ell + F_{np;\mu}}
-  \prod_{q=1}^{p-1} \p{\ell-q+F_{np;\mu}} \\
-  &= { m \choose p + 1 } \prod_{q=0}^p \p{m-p-q+F_{np;\mu}}.
-\end{align}
-We can further simplify
-\begin{align}
-  \prod_{q=0}^p \p{m-p-q+F_{np;\mu}}
-  = \prod_{q=0}^p \p{m-n-q+2\mu S_\z}
-  = \sum_{q=0}^{p+1} \p{-1}^{p+1-q}
-  { p+1 \brack q } \p{m-n+2\mu S_\z}^q,
-\end{align}
-where ${ p \brack q }$ is an unsigned Stirling number of the first
-kind, and finally
-\begin{align}
-  \sum_{q=0}^p \p{-1}^{p-q} { p \brack q } \p{m-n+2\mu S_\z}^q
-  &= \sum_{q=0}^p \p{-1}^{p-q} { p \brack q } \sum_{\ell=0}^q
-  { q \choose \ell } \p{m-n}^{q-\ell} \p{2\mu S_\z}^\ell \\
-  &= \sum_{\ell=0}^p 2^\ell \sum_{q=\ell}^p \p{-1}^{p-q}
-  { p \brack q } { q \choose \ell } \p{m-n}^{q-\ell} \p{\mu S_\z}^\ell.
-\end{align}
-Putting everything together, we finally have
-\begin{align}
-  E_{mn;\mu}^{(p-1)}
-  = p! { m \choose p } { n \choose p }
-  S_\bmu^{n-p} S_\mu^{m-p}
-  \sum_{\ell=0}^p \epsilon_{mn}^{p\ell} \p{\mu S_\z}^\ell,
-\end{align}
-with
-\begin{align}
-  \epsilon_{mn}^{p\ell}
-  \equiv 2^\ell \sum_{q=\ell}^p \p{-1}^{p-q}
-  { p \brack q } { q \choose \ell } \p{m-n}^{q-\ell},
-\end{align}
-where in this final form $E_{mn;\mu}^{(0)} = D_{mn;\mu}$, which
-together with the expansion for $C_{mn;\mu}$ in \eqref{eq:C_mn_E}
-implies that
-\begin{align}
-  \sp{S_\mu^m, S_\bmu^n}_-
-  = \sum_{p=1}^{\min\set{m,n}}
-  p! { m \choose p } { n \choose p } S_\bmu^{n-p} S_\mu^{m-p}
-  \sum_{\ell=0}^p \epsilon_{mn}^{p\ell} \p{\mu S_\z}^\ell,
-  \label{eq:comm_mu}
-\end{align}
-and
-\begin{align}
-  S_\mu^m S_\bmu^n
-  = \sum_{p=0}^{\min\set{m,n}}
-  p! { m \choose p } { n \choose p } S_\bmu^{n-p} S_\mu^{m-p}
-  \sum_{\ell=0}^p \epsilon_{mn}^{p\ell} \p{\mu S_\z}^\ell.
-  \label{eq:push_mu_bmu}
-\end{align}
-If we wish to order products of collective spin operators with $S_\z$
-in between $S_\bmu$ and $S_\mu$, then
-\begin{align}
-  S_\mu^m S_\bmu^n
-  = \sum_{p=0}^{\min\set{m,n}} p! { m \choose p } { n \choose p }
-  S_\bmu^{n-p} Z_{mn;\bmu}^{(p)} S_\mu^{m-p},
-\end{align}
-where
-\begin{align}
-  Z_{mn;\bmu}^{(p)}
-  \equiv \sum_{\ell=0}^p \epsilon_{mn}^{p\ell}
-  \p{-\sp{m-p} + \mu S_\z}^\ell
-  = \sum_{q=0}^p \zeta_{mn}^{pq} \p{\bmu S_\z}^q,
-  \label{eq:Z_mnp}
-\end{align}
-with
-\begin{align}
-  \zeta_{mn}^{pq}
-  \equiv \sum_{\ell=q}^p \epsilon_{mn}^{p\ell}
-  { \ell \choose q } \p{-1}^\ell \p{m-p}^{\ell-q}
-  = \p{-1}^p 2^q \sum_{s=q}^p
-  { p \brack s } { s \choose q } \p{m+n-2p}^{s-q}.
-  \label{eq:zeta_mnpq}
-\end{align}
-Here ${ p \brack s }$ is an unsigned Stirling number of the first
-kind.
-
-\section{Product of arbitrary ordered collective spin operators}
-\label{sec:general_product}
-
-The most general product of collective spin operators that we need to
-compute is
-\begin{align}
-  \S^{pqr}_{\ell mn;\mu}
-  = S_\mu^p \p{\mu S_\z}^q S_\bmu^r
-  S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n
-  = \sum_{k=0}^{\min\set{r,\ell}} k! { r \choose k } { \ell \choose k }
-  S_\mu^{p+\ell-k} \tilde Z_{qr\ell m;\mu}^{(k)} S_\bmu^{r+n-k},
-  \label{eq:general_product}
-\end{align}
-where
-\begin{align}
-  \tilde Z_{qr\ell m;\mu}^{(k)}
-  &\equiv \p{\ell-k+\mu S_\z}^q
-  Z_{r\ell;\mu}^{(k)} \p{r-k+\mu S_\z}^m \\
-  &= \sum_{a=0}^k \zeta_{r\ell}^{ka}
-  \sum_{b=0}^q \p{\ell-k}^{q-b} { q \choose b }
-  \sum_{c=0}^m \p{r-k}^{m-c} { m \choose c }
-  \p{\mu S_\z}^{a+b+c},
-\end{align}
-is defined in terms of $Z_{r\ell;\mu}^{(k)}$ and $\zeta_{r\ell}^{ka}$
-as respectively given in \eqref{eq:Z_mnp} and \eqref{eq:zeta_mnpq}.
-The (anti-)commutator of two ordered products of collective spin
-operators is then simply
-\begin{align}
-  \sp{S_\mu^p \p{\mu S_\z}^q S_\bmu^r,
-    S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_\pm
-  = \S^{pqr}_{\ell mn;\mu} \pm \S^{\ell mn}_{pqr;\mu}.
-\end{align}
-
-\section{Sandwich identities for single-spin decoherence calculations}
-\label{sec:sandwich_single}
-
-In this section we derive several identities that will be necessary
-for computing the effects of single-spin decoherence on ordered
-products of collective spin operators, i.e.~on operators of the form
-$S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$.  These identities all involve
-sandwiching a collective spin operator between operators that act on
-individual spins only, and summing over all individual spin indices.
-Our general strategy will be to use commutation relations to push
-single-spin operators together, and then evaluate the sum to arrive at
-an expression involving only collective spin operators.
-
-We first compute sums of single-spin operators sandwiching
-$\p{\mu S_\z}^m$, when necessary making use of the identity in
-\eqref{eq:push_z_mu_Ss}.  Up to Hermitian conjugation, the unique
-cases are, for $S\equiv N/2$ and $\mu,\nu\in\set{+1,-1}$,
-\begin{align}
-  \sum_j s_\z^{(j)} \p{\mu S_\z}^m s_\z^{(j)}
-  &= \sum_j s_\z^{(j)} s_\z^{(j)} \p{\mu S_\z}^m
-  = \f14 \sum_j \1_j \p{\mu S_\z}^m
-  = \f12 S \p{\mu S_\z}^m, \\
-  \sum_j s_\z^{(j)} \p{\mu S_\z}^m s_\nu^{(j)}
-  &= \p{\mu S_\z}^m \sum_j s_\z^{(j)} s_\nu^{(j)}
-  = \f12 \p{\mu S_\z}^m \nu S_\nu
-  = \f12 \nu S_\nu \p{\mu\nu+\mu S_\z}^m, \\
-  \sum_j s_\nu^{(j)} \p{\mu S_\z}^m s_\nu^{(j)}
-  &= \sum_j s_\nu^{(j)} s_\nu^{(j)} \p{\mu\nu+\mu S_\z}^m
-  = 0, \\
-  \sum_j s_\bnu^{(j)} \p{\mu S_\z}^m s_\nu^{(j)}
-  &= \sum_j s_\bnu^{(j)} s_\nu^{(j)} \p{\mu\nu+\mu S_\z}^m
-  = \p{S-\nu S_\z} \p{\mu\nu+\mu S_\z}^m.
-\end{align}
-We are now equipped to derive similar identities for more general
-collective spin operators.  Making heavy use of identities
-\eqref{eq:push_z_mu_sS} and \eqref{eq:push_mu_Ss} to push single-spin
-operators through transverse collective-spin operators, we again work
-through all combinations that are unique up to Hermitian conjugation,
-finding
-\begin{align}
-  \sum_j s_\z^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\z^{(j)}
-  &= \f12 \p{S-\ell-n} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n
-  + \ell n S_\mu^{\ell-1} \p{S+\mu S_\z}
-  \p{-1+\mu S_\z}^m S_\bmu^{n-1},
-  \label{eq:san_z_z} \allowdisplaybreaks \\
-  \sum_j s_\z^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\mu^{(j)}
-  &= \f12 \mu S_\mu^{\ell+1} \p{1+\mu S_\z}^m S_\bmu^n
-  - \mu n \p{S-\ell-\f12\sp{n-1}} S_\mu^\ell
-  \p{\mu S_\z}^m S_\bmu^{n-1} \notag \\
-  &\qquad - \mu\ell n\p{n-1} S_\mu^{\ell-1}
-  \p{S+\mu S_\z} \p{-1+\mu S_\z}^m S_\bmu^{n-2},
-  \label{eq:san_z_mu} \allowdisplaybreaks \\
-  \sum_j s_\z^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\bmu^{(j)}
-  &= -\f12 \mu S_\mu^\ell \p{\mu S_\z}^m S_\bmu^{n+1}
-  + \mu \ell S_\mu^{\ell-1} \p{S+\mu S_\z} \p{-1+\mu S_\z}^m S_\bmu^n,
-  \label{eq:san_z_bmu} \allowdisplaybreaks \\
-  \sum_j s_\mu^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\mu^{(j)}
-  &= n S_\mu^{\ell+1} \p{\mu S_\z}^m S_\bmu^{n-1}
-  - n\p{n-1} S_\mu^\ell \p{S+\mu S_\z} \p{-1+\mu S_\z}^m S_\bmu^{n-2},
-  \label{eq:san_mu_mu} \allowdisplaybreaks \\
-  \sum_j s_\mu^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\bmu^{(j)}
-  &= S_\mu^\ell \p{S+\mu S_\z}\p{-1+\mu S_\z}^m S_\bmu^n,
-  \label{eq:san_mu_bmu} \allowdisplaybreaks \\
-  \sum_j s_\bmu^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\mu^{(j)}
-  &= S_\mu^\ell \p{S - \ell - n - \mu S_\z}
-  \p{1+\mu S_\z}^m S_\bmu^n
-  + \ell n \p{2S - \ell - n + 2}
-  S_\mu^{\ell-1} \p{\mu S_\z}^m S_\bmu^{n-1} \notag \\
-  &\qquad + \ell n \p{\ell-1} \p{n-1} S_\mu^{\ell-2} \p{S+\mu S_\z}
-  \p{-1+\mu S_\z}^m S_\bmu^{n-2}.
-  \label{eq:san_bmu_mu}
-\end{align}
-
-\section{Uncorrelated, permutationally-symmetric single-spin
-  decoherence}
-\label{sec:decoherence_single}
-
-In this section we work out the effects of permutationally-symmetric
-decoherence of individual spins on collective spin operators of the
-form $S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$.  For compactness, we define
-\begin{align}
-  \D\p{g} \O
-  \equiv \D\p{\set{g^{(j)}:j=1,2,\cdots,N}} \O
-  = \sum_j\p{{g^{(j)}}^\dag \O g^{(j)}
-    - \f12\sp{{g^{(j)}}^\dag g^{(j)}, \O}_+},
-\end{align}
-where $g$ is an operator that acts on a single spin, $g^{(j)}$ is an
-operator that acts with $g$ on spin $j$ and trivially on all other
-spins, and $N$ is the total number of spins.
-
-\subsection{Decay-type decoherence}
-\label{sec:decay_single}
-
-The effect of decoherence via uncorrelated decay ($\mu=-1$) or
-excitation ($\mu=1$) of individual spins is described by
-\begin{align}
-  \D\p{s_\mu} \O
-  = \sum_j\p{s_\bmu^{(j)} \O s_\mu^{(j)}
-    - \f12\sp{s_\bmu^{(j)} s_\mu^{(j)},\O}_+}
-  = \sum_j s_\bmu^{(j)} \O s_\mu^{(j)}
-  - S \O + \f{\mu}{2} \sp{S_\z, \O}_+.
-\end{align}
-In order to determine the effect of this decoherence on general
-collective spin operators, we expand the anti-commutator
-\begin{align}
-  \sp{S_\z, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
-  = S_\z S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n
-  + S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\z
-  = \mu S_\mu^\ell\p{\ell+n+2\mu S_\z} \p{\mu S_\z}^m S_\bmu^n,
-\end{align}
-which implies, using \eqref{eq:san_mu_bmu},
-\begin{align}
-  \D\p{s_\bmu} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
-  = S_\mu^\ell \p{S+\mu S_\z}\p{-1+\mu S_\z}^m S_\bmu^n
-  - S_\mu^\ell\sp{S + \f12\p{\ell+n} + \mu S_\z}
-  \p{\mu S_\z}^m S_\bmu^n,
-  \label{eq:decay_diff}
-\end{align}
-and, using \eqref{eq:san_bmu_mu},
-\begin{align}
-  \D\p{s_\mu} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
-  &= S_\mu^\ell \p{S - \ell - n - \mu S_\z} \p{1+\mu S_\z}^m S_\bmu^n
-  - S_\mu^\ell\sp{S - \f12\p{\ell+n} - \mu S_\z}
-  \p{\mu S_\z}^m S_\bmu^n \notag \\
-  &\qquad + \ell n \p{2S - \ell - n + 2}
-  S_\mu^{\ell-1} \p{\mu S_\z}^m S_\bmu^{n-1} \notag \\
-  &\qquad + \ell n \p{\ell-1} \p{n-1} S_\mu^{\ell-2} \p{S + \mu S_\z}
-  \p{-1+\mu S_\z}^m S_\bmu^{n-2}.
-  \label{eq:decay_same}
-\end{align}
-Decoherence via jump operators $s_\bmu^{(j)}$ only couples operators
-$S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$ to operators
-$S_\mu^\ell \p{\mu S_\z}^{m'} S_\bmu^n$ with $m'\le m$.  Decoherence
-via jump operators $s_\mu^{(j)}$, meanwhile, makes operators
-$S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$ ``grow'' in $m$ through the last
-term in \eqref{eq:decay_same}, although the sum $\ell+m+n$ does not
-grow.
-
-\subsection{Dephasing}
-\label{sec:dephasing_single}
-
-The effect of decoherence via single-spin dephasing is described by
-\begin{align}
-  \D\p{s_\z} \O
-  = \sum_j\p{s_\z^{(j)} \O s_\z^{(j)}
-    - \f12\sp{s_\z^{(j)} s_\z^{(j)},\O}_+}
-  = \sum_j s_\z^{(j)} \O s_\z^{(j)} - \f12 S \O.
-\end{align}
-From \eqref{eq:san_z_z}, we then have
-\begin{align}
-  \D\p{s_\z} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
-  = -\f12\p{\ell+n} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n
-  + \ell n S_\mu^{\ell-1} \p{S + \mu S_\z}
-  \p{-1 + \mu S_\z}^m S_\bmu^{n-1}.
-\end{align}
-Decoherence via single-spin dephasing makes operators
-$S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$ ``grow'' in $m$, although the sum
-$\ell+m+n$ does not grow.
-
-\subsection{The general case}
-\label{sec:general_single}
-
-The most general type of single-spin decoherence is described by
-\begin{align}
-  \D\p{g} \O
-  = \sum_j\p{{g^{(j)}}^\dag \O g^{(j)}
-    - \f12\sp{{g^{(j)}}^\dag g^{(j)}, \O}_+},
-  &&
-  g \equiv g_\z s_\z + g_+ s_+ + g_- s_-.
-  \label{eq:D_general_single}
-\end{align}
-To simplify \eqref{eq:D_general_single}, we expand
-\begin{align}
-  g^\dag \O g
-  = \abs{g_\z}^2 s_\z \O s_\z
-  + \sum_\mu \p{\abs{g_\mu}^2 s_\bmu \O s_\mu
-    + g_\bmu^* g_\mu s_\mu \O s_\mu
-    + g_\z^* g_\mu s_\z \O s_\mu
-    + g_\bmu^* g_\z s_\mu \O s_\z},
-\end{align}
-and
-\begin{align}
-  g^\dag g
-  = \f14 \abs{g_\z}^2
-  + \f12 \sum_\mu \sp{\abs{g_\mu}^2 \p{1-2\mu s_\z}
-    + \mu \p{g_\z^*g_\mu - g_\bmu^*g_\z} s_\mu},
-\end{align}
-which implies
-\begin{align}
-  \D\p{g} \O
-  &= \sum_{X\in\set{\z,+,-}} \abs{g_X}^2 \D\p{s_X} \O
-  + \sum_{\mu,j}
-  \p{g_\bmu^* g_\mu s_\mu^{(j)} \O s_\mu^{(j)}
-    + g_\z^* g_\mu s_\z^{(j)} \O s_\mu^{(j)}
-    + g_\bmu^* g_\z s_\mu^{(j)} \O s_\z^{(j)}}
-  \notag \\
-  &\qquad -\f14 \sum_\mu \mu
-  \p{g_\z^*g_\mu - g_\bmu^*g_\z} \sp{S_\mu, \O}_+.
-\end{align}
-In order to compute the effect of this decoherence on general
-collective spin operators, we expand the anti-commutator
-\begin{align}
-  \sp{S_\mu, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
-  = S_\mu^{\ell+1} \sp{\p{\mu S_\z}^m+\p{1+\mu S_\z}^m} S_\bmu^n
-  - n S_\mu^\ell \p{n-1+2\mu S_\z} \p{\mu S_\z}^m S_\bmu^{n-1}.
-  \label{eq:S_mu_acomm}
-\end{align}
-Recognizing a resemblance between terms in \eqref{eq:S_mu_acomm} and
-\eqref{eq:san_z_mu}, we collect terms to simplify
-\begin{align}
-  \sum_j s_\z^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\mu^{(j)}
-  - \f14 \mu \sp{S_\mu, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
-  = K_{\ell mn;\mu} + L_{\ell mn;\mu}
-  \label{eq:dec_z_mu}
-\end{align}
-and likewise
-\begin{align}
-  \sum_j s_\mu^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\z^{(j)}
-  + \f14 \mu \sp{S_\mu, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
-  = K_{\ell mn;\mu} + M_{\ell mn;\mu}
-  \label{eq:dec_mu_z}
-\end{align}
-with
-\begin{align}
-  K_{\ell mn;\mu}
-  &\equiv \f14 \mu S_\mu^{\ell+1}
-  \sp{\p{1+\mu S_\z}^m-\p{\mu S_\z}^m} S_\bmu^n, \\
-  L_{\ell mn;\mu}
-  &\equiv -\mu n S_\mu^\ell \sp{S-\ell-\f34\p{n-1}-\f12\mu S_\z}
-  \p{\mu S_\z}^m S_\bmu^{n-1}
-  - \mu\ell n\p{n-1} S_\mu^{\ell-1}
-  \p{S+\mu S_\z} \p{-1+\mu S_\z}^m S_\bmu^{n-2}, \\
-  M_{\ell mn;\mu}
-  &\equiv \mu n S_\mu^\ell \sp{\p{S+\mu S_\z}\p{-1+\mu S_\z}^m
-    - \f12\p{\f12\sp{n-1}+\mu S_\z}\p{\mu S_\z}^m} S_\bmu^{n-1}.
-\end{align}
-Defining for completion
-\begin{align}
-  P_{\ell mn;\mu}
-  &\equiv \sum_j s_\mu^{(j)} S_\mu^\ell
-  \p{\mu S_\z}^m S_\bmu^n s_\mu^{(j)}
-  = n S_\mu^{\ell+1} \p{\mu S_\z}^m S_\bmu^{n-1}
-  - n\p{n-1} S_\mu^\ell \p{S+\mu S_\z} \p{-1+\mu S_\z}^m S_\bmu^{n-2},
-\end{align}
-and
-\begin{align}
-  Q_{\ell mn;\mu}^{(g)}
-  \equiv g_\bmu^* g_\mu P_{\ell mn;\mu}
-  + \p{g_\z^* g_\mu + g_\bmu^* g_\z}
-  K_{\ell mn;\mu}
-  + g_\z^* g_\mu L_{\ell mn;\mu}
-  + g_\bmu^* g_\z M_{\ell mn;\mu},
-  \label{eq:Q_single}
-\end{align}
-we finally have
-\begin{align}
-  \D\p{g} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
-  = \sum_{X\in\set{\z,+,-}} \abs{g_X}^2
-  \D\p{s_X} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
-  + Q_{\ell mn;\mu}^{(g)} + \sp{Q_{nm\ell;\mu}^{(g)}}^\dag.
-\end{align}
-Note that the sum $\ell+m+n$ for operators
-$S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$ does not grow under this type of
-decoherence.
-
-\section{Sandwich identities for collective-spin decoherence
-  calculations}
-\label{sec:sandwich_collective}
-
-In analogy with the work in Appendix \ref{sec:sandwich_single}, in
-this section we work out sandwich identities necessary for
-collective-spin decoherence calculations.  The simplest cases are
-\begin{align}
-  S_\mu S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\bmu
-  &= S_\mu^{\ell+1} \p{\mu S_\z}^m S_\bmu^{n+1},
-  \allowdisplaybreaks \\
-  S_\mu S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\z
-  &= \mu S_\mu^{\ell+1} \p{n+\mu S_\z} \p{\mu S_\z}^m S_\bmu^n,
-  \allowdisplaybreaks \\
-  S_\z S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\z
-  &= S_\mu^\ell \sp{\ell n + \p{\ell+n} \mu S_\z + \p{\mu S_\z}^2}
-  \p{\mu S_\z}^m S_\bmu^n.
-\end{align}
-With a bit more work, we can also find
-\begin{align}
-  S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\mu
-  &= S_\mu^{\ell+1} \p{1+\mu S_\z}^m S_\bmu^n
-  - n S_\mu^\ell \p{n-1+2\mu S_\z} \p{\mu S_\z}^m S_\bmu^{n-1},
-\end{align}
-which implies
-\begin{align}
-  S_\mu S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\mu
-  &= S_\mu^{\ell+2} \p{1+\mu S_\z}^m S_\bmu^n
-  - n S_\mu^{\ell+1} \p{n-1+2\mu S_\z} \p{\mu S_\z}^m S_\bmu^{n-1},
-  \allowdisplaybreaks \\
-  S_\z S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\mu
-  &= \mu S_\mu^{\ell+1} \p{\ell+1+\mu S_\z} \p{1+\mu S_\z}^m S_\bmu^n
-  \notag \\
-  &\qquad - \mu n S_\mu^\ell
-  \sp{\ell\p{n-1} + \p{2\ell+n-1}\mu S_\z + 2\p{\mu S_\z}^2}
-  \p{\mu S_\z}^m S_\bmu^{n-1}.
-\end{align}
-Finally, we compute
-\begin{align}
-  S_\bmu S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\mu
-  &= \sp{S_\mu^\ell S_\bmu - \ell S_\mu^{\ell-1} \p{\ell-1+2\mu S_\z}}
-  \p{\mu S_\z}^m
-  \sp{S_\mu S_\bmu^n - n \p{n-1+2\mu S_\z} S_\bmu^{n-1}} \notag \\
-  &= S_\mu^\ell S_\bmu \p{\mu S_\z}^m S_\mu S_\bmu^n \notag \\
-  &\qquad - S_\mu^\ell
-  \sp{\ell\p{\ell+1} + n\p{n+1}+2\p{\ell+n}\mu S_\z}
-  \p{1+\mu S_\z}^m S_\bmu^n \notag \\
-  &\qquad + \ell n S_\mu^{\ell-1}
-  \sp{\p{\ell-1}\p{n-1}+2\p{\ell+n-2}\mu S_\z + 4\p{\mu S_\z}^2}
-  \p{\mu S_\z}^m S_\bmu^{n-1},
-\end{align}
-where
-\begin{multline}
-  S_\bmu \p{\mu S_\z}^m S_\mu
-  = S_\bmu S_\mu \p{1+\mu S_\z}^m
-  = \p{S_\mu S_\bmu - 2\mu S_\z} \p{1+\mu S_\z}^m
-  = S_\mu \p{2+\mu S_\z}^m S_\bmu - 2\mu S_\z \p{1+\mu S_\z}^m,
-\end{multline}
-so
-\begin{align}
-  S_\bmu S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\mu
-  &= S_\mu^{\ell+1} \p{2+\mu S_\z}^m S_\bmu^{n+1} \notag \\
-  &\qquad - S_\mu^\ell
-  \sp{\ell\p{\ell+1} + n\p{n+1}+2\p{\ell+n+1}\mu S_\z}
-  \p{1+\mu S_\z}^m S_\bmu^n \notag \\
-  &\qquad + \ell n S_\mu^{\ell-1}
-  \sp{\p{\ell-1}\p{n-1}+2\p{\ell+n-2}\mu S_\z + 4\p{\mu S_\z}^2}
-  \p{\mu S_\z}^m S_\bmu^{n-1}.
-\end{align}
-
-\section{Collective spin decoherence}
-\label{sec:decoherence_collective}
-
-In this section we work out the effects of collective decoherence on
-general collective spin operators.  For shorthand, we define
-\begin{align}
-  \D\p{G} \O
-  \equiv \D\p{\set{G}} \O
-  = G^\dag \O G - \f12\sp{G^\dag G, \O}_+,
-\end{align}
-where $G$ is a collective spin jump operator.
-
-\subsection{Decay-type decoherence and dephasing}
-\label{sec:decay_dephasing_collective}
-
-Making use of the results in Appendix \ref{sec:sandwich_collective},
-we find that the effects of collective decay-type decoherence on
-general collective spin operators are given by
-\begin{align}
-  \D\p{S_\bmu} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
-  &= -S_\mu^{\ell+1} \sp{\p{1+\mu S_\z}^m - \p{\mu S_\z}^m} S_\bmu^{n+1}
-  \notag \\
-  &\qquad + \f12 S_\mu^\ell \sp{\ell\p{\ell-1} + n\p{n-1}
-    + 2\p{\ell+n}\mu S_\z} \p{\mu S_\z}^m S_\bmu^n,
-\end{align}
-and
-\begin{align}
-  \D\p{S_\mu} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
-  &= S_\mu^{\ell+1} \sp{\p{2+\mu S_\z}^m-\p{1+\mu S_\z}^m} S_\bmu^{n+1}
-  \notag \\
-  &\qquad - S_\mu^\ell
-  \sp{\ell\p{\ell+1} + n\p{n+1}+2\p{\ell+n+1}\mu S_\z}
-  \p{1+\mu S_\z}^m S_\bmu^n \notag \\
-  &\qquad + \f12 S_\mu^\ell
-  \sp{\ell\p{\ell+1} + n\p{n+1}+2\p{\ell+n+2}\mu S_\z}
-  \p{\mu S_\z}^m S_\bmu^n \notag \\
-  &\qquad + \ell n S_\mu^{\ell-1}
-  \sp{\p{\ell-1}\p{n-1}+2\p{\ell+n-2}\mu S_\z + 4\p{\mu S_\z}^2}
-  \p{\mu S_\z}^m S_\bmu^{n-1}.
-\end{align}
-Similarly, the effect of collective dephasing is given by
-\begin{align}
-  \D\p{S_\z} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
-  = -\f12 \p{\ell-n}^2 S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n.
-\end{align}
-
-\subsection{The general case}
-\label{sec:general_collective}
-
-More generally, we consider jump operators of the form
-\begin{align}
-  G \equiv G_\z S_\z + G_+ S_+ + G_- S_-,
-\end{align}
-whose decoherence effects are determined by
-\begin{align}
-  G^\dag \O G
-  = \abs{G_\z}^2 S_\z \O S_\z
-  + \sum_\mu \p{\abs{G_\mu}^2 S_\bmu \O S_\mu
-    + G_\bmu^* G_\mu S_\mu \O S_\mu
-    + G_\z^* G_\mu S_\z \O S_\mu
-    + G_\bmu^* G_\z S_\mu \O S_\z},
-\end{align}
-and
-\begin{align}
-  G^\dag G
-  = \abs{G_\z}^2 S_\z^2
-  + \sum_\mu \p{\abs{G_\mu}^2 S_\bmu S_\mu
-    + G_\z^*G_\mu S_\z S_\mu
-    + G_\bmu^* G_\z S_\mu S_\z
-    + G_\bmu^* G_\mu S_\mu^2},
-\end{align}
-which implies
-\begin{align}
-  \D\p{G} \O
-  &= \sum_{X\in\set{\z,+,-}} \abs{G_X}^2 \D\p{S_X} \O
-  + \sum_\mu \p{G_\bmu^* G_\mu S_\mu \O S_\mu
-    + G_\z^* G_\mu S_\z \O S_\mu
-    + G_\bmu^* G_\z S_\mu \O S_\z}
-  \notag \\
-  &\qquad -\f12 \sum_\mu\p{G_\bmu^* G_\mu \sp{S_\mu^2, \O}_+
-    + G_\z^*G_\mu \sp{S_\z S_\mu, \O}_+
-    + G_\bmu^* G_\z \sp{S_\mu S_\z, \O}_+}.
-\end{align}
-In order to compute the effect of this decoherence on general
-collective spin operators, we expand the anti-commutators
-\begin{align}
-  \sp{S_\mu^2, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
-  &= S_\mu^{\ell+2} \sp{\p{2+\mu S_\z}^m+\p{\mu S_\z}^m} S_\bmu^n
-  - 2n S_\mu^{\ell+1} \p{n+2\mu S_\z} \p{1+\mu S_\z}^m S_\bmu^{n-1}
-  \notag \\
-  &\qquad + n\p{n-1} S_\mu^\ell \sp{\p{n-1}\p{n-2}
-    + 2\p{2n-3}\mu S_\z + 4\p{\mu S_\z}^2} \p{\mu S_\z}^m S_\bmu^{n-2},
-  \allowdisplaybreaks \\
-  \sp{S_\z S_\mu, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
-  &= \mu S_\mu^{\ell+1} \sp{\p{\ell+1+\mu S_\z}\p{\mu S_\z}^m
-    + \p{n+1+\mu S_\z} \p{1+\mu S_\z}^m } S_\bmu^n \notag \\
-  &\qquad - \mu n S_\mu^\ell \sp{n \p{n-1}
-    + \p{3n-1}\mu S_\z + 2\p{\mu S_\z}^2} \p{\mu S_\z}^m S_\bmu^{n-1},
-  \allowdisplaybreaks \\
-  \sp{S_\mu S_\z, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
-  &= \mu S_\mu^{\ell+1} \sp{\p{\ell+\mu S_\z}\p{\mu S_\z}^m
-    + \p{n+\mu S_\z} \p{1+\mu S_\z}^m} S_\bmu^n \notag \\
-  &\qquad - \mu n S_\mu^\ell \sp{\p{n-1}^2
-    + 3\p{n-1}\mu S_\z + 2\p{\mu S_\z}^2} \p{\mu S_\z}^m S_\bmu^{n-1}.
-\end{align}
-Collecting terms and defining
-\begin{align}
-  G_{\z,\mu}^{(\pm)}
-  &\equiv \f12\p{G_\z^* G_\mu \pm G_\bmu^* G_\z},
-  \allowdisplaybreaks \\
-  \tilde L_{\ell mn;\mu}^{(G)}
-  &\equiv \mu \sp{\p{\ell-n+\f12} G_{\z,\mu}^{(+)}
-    + \p{\ell+\f12} G_{\z,\mu}^{(-)}}
-  S_\mu^{\ell+1} \p{1+\mu S_\z}^m S_\bmu^n \notag \\
-  &\qquad -\mu \sp{\p{\ell-n+\f12} G_{\z,\mu}^{(+)}
-    + \p{n+\f12} G_{\z,\mu}^{(-)}}
-  S_\mu^{\ell+1} \p{\mu S_\z}^m S_\bmu^n \notag \\
-  &\qquad + \mu G_{\z,\mu}^{(-)}
-  S_\mu^{\ell+1} \mu S_\z
-  \sp{\p{1+\mu S_\z}^m - \p{\mu S_\z}^m} S_\bmu^n,
-  \allowdisplaybreaks \\
-  \tilde M_{\ell mn;\mu}^{(G)}
-  &= -\mu n\p{n-1} \sp{\p{\ell-n+\f12} G_{\z,\mu}^{(+)}
-    + \p{\ell-\f12} G_{\z,\mu}^{(-)}}
-  S_\mu^\ell \p{\mu S_\z}^m S_\bmu^{n-1} \notag \\
-  &\qquad - 2\mu n \sp{\p{\ell-n+\f12} G_{\z,\mu}^{(+)}
-    + \p{\ell+\f12n-1} G_{\z,\mu}^{(-)}}
-  S_\mu^\ell \p{\mu S_\z}^{m+1} S_\bmu^{n-1} \notag \\
-  &\qquad - 2\mu n G_{\z,\mu}^{(-)}
-  S_\mu^\ell \p{\mu S_\z}^{m+2} S_\bmu^{n-1},
-  \allowdisplaybreaks \\
-  \tilde P_{\ell mn;\mu}
-  &\equiv -\f12 S_\mu^{\ell+2}
-  \sp{\p{2+\mu S_\z}^m - 2\p{1+\mu S_\z}^m + \p{\mu S_\z}^m}
-  S_\bmu^n \notag \\
-  &\qquad + n S_\mu^{\ell+1} \sp{\p{n+2\mu S_\z} \p{1+\mu S_\z}^m
-    - \p{n-1+2\mu S_\z} \p{\mu S_\z}^m}
-  S_\bmu^{n-1} \notag \\
-  &\qquad -n\p{n-1} S_\mu^\ell
-  \sp{\f12\p{n-1}\p{n-2} + \p{2n-3}\mu S_\z + 2\p{\mu S_\z}^2}
-  \p{\mu S_\z}^m S_\bmu^{n-2},
-  \allowdisplaybreaks \\
-  \tilde Q_{\ell mn;\mu}^{(G)}
-  &\equiv G_\bmu^* G_\mu \tilde P_{\ell mn;\mu}
-  + \tilde L_{\ell mn;\mu}^{(G)}
-  + \tilde M_{\ell mn;\mu}^{(G)},
-\end{align}
-we then have
-\begin{align}
-  \D\p{G} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
-  = \sum_{X\in\set{\z,+,-}} \abs{G_X}^2
-  \D\p{S_X} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
-  + \tilde Q_{\ell mn;\mu}^{(G)}
-  + \sp{\tilde Q_{nm\ell;\mu}^{(G)}}^\dag.
-\end{align}
-Note that the sum $\ell+m+n$ for operators
-$S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$ grows by one if $G_\mu\ne0$ or
-$G_\bmu\ne0$, and does not grow otherwise.
-
-\section{Initial conditions}
-\label{sec:initial_conditions}
-
-Here we compute the expectation values of collective spin operators
-with respect to spin-polarized (Gaussian) states.  These states are
-parameterized by polar and azimuthal angles $\theta\in[0,\pi)$,
-$\phi\in[0,2\pi)$, and lie within the Dicke manifold spanned by states
-$\ket{k}\propto S_+^{S+k}\ket{\dn}^{\otimes N}$ with $S\equiv N/2$ and
-$S_\z\ket{k}=k\ket{k}$:
-\begin{align}
-  \ket{\theta,\phi}
-  \equiv \sp{\cos\p{\theta/2} e^{-i\phi/2} \ket\up
-    + \sin\p{\theta/2} e^{i\phi/2} \ket\dn}^{\otimes N}
-  = \sum_{k=-S}^S { N \choose S+k }^{1/2}
-  \cos\p{\theta/2}^{S+k} \sin\p{\theta/2}^{S-k} e^{-ik\phi} \ket{k}.
-\end{align}
-We can likewise expand, within the Dicke manifold,
-\begin{align}
-  S_\z = \sum_{k=-S}^S k \op{k},
-  &&
-  S_\mu = \sum_{k=-S+\delta_{\mu,-1}}^{S-\delta_{\mu,1}}
-  g_\mu\p{k} \op{k+\mu}{k}
-  = \sum_{k=-S+\delta_{\bmu,-1}}^{S-\delta_{\bmu,1}}
-  g_\bmu\p{k} \op{k}{k+\bmu},
-\end{align}
-where $\bmu\equiv-\mu\in\set{+1,-1}$ and
-\begin{align}
-  g_\mu\p{k} \equiv \sqrt{\p{S-\mu k}\p{S+\mu k+1}},
-\end{align}
-which implies
-\begin{align}
-  S_\mu^\ell \p{\mu S_\z^m} S_\bmu^n
-  &= \sum_{k=-S+\delta_{\mu,-1}\max\set{\ell,n}}
-  ^{S-\delta_{\mu,1}\max\set{\ell,n}} \p{\mu k}^m
-  \sp{\prod_{p=0}^{\ell-1} g_\mu\p{k+\mu p}}
-  \sp{\prod_{q=0}^{n-1} g_\mu\p{k+\mu q}}
-  \op{k+\mu\ell}{k+\mu n} \\
-  &= \sum_{\mu k=-\mu S-\delta_{\mu,-1}\max\set{\ell,n}}
-  ^{\mu S-\delta_{\mu,1}\max\set{\ell,n}} \p{\mu k}^m
-  \f{\p{S-\mu k}!}{\p{S+\mu k}!}
-  \sp{\f{\p{S+\mu k+\ell}!}{\p{S-\mu k-\ell}!}
-    \f{\p{S+\mu k+n}!}{\p{S-\mu k-n}!}}^{1/2}
-  \op{k+\mu\ell}{k+\mu n} \\
-  &= \sum_{k=-S}^{S-\max\set{\ell,n}} k^m
-  \f{\p{S-k}!}{\p{S+k}!}
-  \sp{\f{\p{S+k+\ell}!}{\p{S-k-\ell}!}
-    \f{\p{S+k+n}!}{\p{S-k-n}!}}^{1/2}
-  \op{\mu\p{k+\ell}}{\mu\p{k+n}}.
-\end{align}
-This expansion allows us to compute the expectation value
-\begin{align}
-  \bk{\theta,\phi|S_\mu^\ell \p{\mu S_\z^m} S_\bmu^n|\theta,\phi}
-  &= e^{i\phi \mu\p{\ell-n}} N! \sum_{k=-S}^{S-\max\set{\ell,n}}
-  \f{k^m \p{S-k}! f_{\mu\ell n}\p{k,\theta}}
-  {\p{S+k}! \p{S-k-\ell}! \p{S-k-n}!} \\
-  &= e^{i\phi \mu\p{\ell-n}} \p{-1}^m N! \sum_{k=0}^{N-\max\set{\ell,n}}
-  \f{\p{S-k}^m \p{N-k}! \tilde f_{\mu\ell n}\p{k,\theta}}
-  {k! \p{N-k-\ell}! \p{N-k-n}!}
-\end{align}
-where
-\begin{align}
-  f_{\mu\ell n}\p{k,\theta}
-  \equiv \cos\p{\theta/2}^{N+\mu\p{2k+\ell+n}}
-  \sin\p{\theta/2}^{N-\mu\p{2k+\ell+n}},
-\end{align}
-\begin{align}
-  \tilde f_{\mu\ell n}\p{k,\theta}
-  \equiv f_{\mu\ell n}\p{k-S,\theta}
-  = \cos\p{\theta/2}^{2N\delta_{\mu,-1}+\mu\p{2k+\ell+n}}
-  \sin\p{\theta/2}^{2N\delta_{\mu,1}-\mu\p{2k+\ell+n}}.
-\end{align}
-Defining the states
-\begin{align}
-  \ket{+\Z} \equiv \ket{0,0} = \ket\up^{\otimes N}, &&
-  \ket{-\Z} \equiv \ket{\pi,0} = \ket\dn^{\otimes N}, &&
-  \ket\X \equiv \ket{\pi/2,0}
-  = \p{\f{\ket\up+\ket\dn}{\sqrt2}}^{\otimes N},
-\end{align}
-some particular expectation values of interest are
-\begin{align}
-  \bk{\nu\Z|S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n|\nu\Z}
-  = \delta_{\ell n} \times
-  \begin{cases}
-    \p{S-n}^m \f{N! n!}{\p{N-n}!} & \mu = \nu, \\
-    \delta_{n,0} \p{-S}^m & \mu \ne \nu,
-  \end{cases},
-\end{align}
-and
-\begin{align}
-  \bk{\X|S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n|\X}
-  = \p{-1}^m \f{N!}{2^N} \sum_{k=0}^{N-\max\set{\ell,n}}
-  \f{\p{S-k}^m \p{N-k}!}{k!\p{N-k-\ell}!\p{N-k-n}!}.
-\end{align}
-
-\section{Computing correlators with the truncated short-time (TST)
-  expansion}
-\label{sec:tutorial}
-
-Here, we provide a pedagogical tutorial for computing correlators
-using the truncated short-time TST expansion.  For concreteness, we
-nominally consider $N$ spins evolving under the one-axis twisting
-(OAT) Hamiltonian
-\begin{align}
-  H_{\t{OAT}} = \chi S_\z^2,
-\end{align}
-additionally subject to spontaneous single-spin decay at rate
-$\gamma_-$, with jump operators $\A_-=\set{s_-^{(j)}:j=1,2,\cdots,N}$.
-The equation of motion for a Heisenberg operator
-$S_+^\ell S_\z^m S_-^n$ is
-\begin{align}
-  \f{d}{dt} \p{S_+^\ell S_\z^m S_-^n}
-  = i\chi\sp{S_\z^2, S_+^\ell S_\z^m S_-^n}_-
-  + \gamma_- \D\p{\A_-} \p{S_+^\ell S_\z^m S_-^n},
-  \label{eq:tutorial_EOM}
-\end{align}
-where we can use the results in appendices \ref{sec:general_product}
-and \ref{sec:decoherence_single} respectively to evaluate the
-commutator $\sp{S_\z^2, S_+^\ell S_\z^m S_-^n}_-$ and dissipator
-$\D\p{\A_-} \p{S_+^\ell S_\z^m S_-^n}$ in \eqref{eq:tutorial_EOM} (see
-\eqref{eq:general_product} and \eqref{eq:decay_diff} in particular),
-giving us
-\begin{align}
-  \f{d}{dt} \p{S_+^\ell S_\z^m S_-^n} =
-  i\chi \p{\ell-n} S_+^\ell \p{\ell+n+2S_\z} S_\z^m S_-^n
-  + \gamma_- S_+^\ell \sp{\p{S+S_\z}\p{-1+S_\z}^m
-    - \p{S+\f{\ell+n}{2}+S_\z} S_\z^m} S_-^n.
-  \label{eq:tutorial_EOM_expanded}
-\end{align}
-In practice, we do not want to keep track of such an expansion by
-hand, especially in the case of e.g.~the two-axis twisting (TAT) and
-twist-and-turn (TNT) models with more general types of decoherence,
-for which the analogue of \eqref{eq:tutorial_EOM_expanded} may take
-several lines just to write out in full.  Defining the operators
-$\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$ with
-$\v m\equiv\p{m_+,m_\z,m_-}$ for shorthand, we note that the vector
-space spanned by $\set{\S_{\v m}}$ is closed under time evolution.  We
-therefore expand
-\begin{align}
-  \f{d}{dt} \S_{\v n} \equiv T \S_{\v n}
-  = \sum_{\v m} \S_{\v m} T_{\v m\v n},
-  \label{eq:tutorial_EOM_general}
-\end{align}
-where $T$ is a superoperator that generates time evolution for
-Heisenberg operators.  In the present example, the matrix elements
-$T_{\v m\v n}\in\C$ of $T$ are defined by
-\eqref{eq:tutorial_EOM_expanded} and \eqref{eq:tutorial_EOM_general}.
-For any Hamiltonian $H$ with decoherence characterized by sets of jump
-operators $\A$ and decoherence rates $\gamma_\A$, the matrix elements
-$T_{\v m\v n}$ are more generally defined by
-\begin{align}
-  T \S_{\v n}
-  = i\sp{H,\S_{\v n}}_- + \sum_\A \gamma_\A \D\p{\A} \S_{\v n}
-  = \sum_{\v m} \S_{\v m} T_{\v m\v n}.
-\end{align}
-The results in Appendices \ref{sec:general_product},
-\ref{sec:decoherence_single}, and \ref{sec:decoherence_collective} can
-be used to write model-agnostic codes that compute matrix elements
-$T_{\v m\v n}$, taking a particular Hamiltonian $H$ and decoherence
-processes $\set{\p{\A,\gamma_\A}}$ as inputs.
-
-In order to compute a quantity such as spin squeezing, we need to
-compute correlators of the form $\bk{\S_{\v n}\p{t}}$, where
-$\S_{\v n}\p{t}$ is the time-evolved Heisenberg operator corresponding
-to $\S_{\v n}$, with $\S_{\v n}\p{0}\equiv\S_{\v n}$.  The order-$M$
-truncated short-time (TST) expansion takes
-\begin{align}
-  \bk{\S_{\v n}\p{t}}
-  \to \sum_{k=0}^M \f{t^k}{k!}
-  \sum_{\v m} \bk{\S_{\v m}\p{0}} T^k_{\v m\v n},
-  \label{eq:tutorial_TST}
-\end{align}
-where $T^k_{\v m\v n}$ are matrix elements of the $k$-th time
-derivative operator, given by
-\begin{align}
-  T^0_{\v m\v n} \equiv
-    \begin{cases}
-      1 & \v m = \v n, \\
-      0 & \t{otherwise}
-    \end{cases},
-  &&
-  T^1_{\v m\v n} \equiv T_{\v m\v n},
-  &&
-  T^{k>1}_{\v m\v n}
-  \equiv \sum_{\v p_1,\v p_2,\cdots,\v p_{k-1}}
-  T_{\v m\v p_{k-1}} \cdots T_{\v p_3\v p_2}
-  T_{\v p_2\v p_1} T_{\v p_1\v n}.
-\end{align}
-Matrix elements $T^k_{\v m\v n}$ and initial-time expectation values
-$\bk{\S_{\v m}\p{0}}$ are thus computed as needed for any particular
-correlator $\bk{\S_{\v n}\p{t}}$ of interest, and combined according
-to \eqref{eq:tutorial_TST}.  Note that initial-time expectation values
-$\bk{\S_{\v m}\p{0}}$ are an input to the TST expansion, and need to
-be computed separately for any given initial state; expectation values
-with respect to spin-polarized (Gaussian) states are provided in
-Appendix \ref{sec:initial_conditions}.  In practice, we further
-collect terms in \eqref{eq:tutorial_TST} to write
-\begin{align}
-  \bk{\S_{\v n}\p{t}} \to \sum_{k=0}^M c_{\v n k} t^k,
-  &&
-  c_{\v n k}
-  \equiv \f1{k!} \sum_{\v m} \bk{\S_{\v m}\p{0}} T^k_{\v m\v n},
-\end{align}
-where $c_{\v n k}$ are time-independent coefficients for the
-polynomial time expansion of $\bk{\S_{\v n}\p{t}}$.  After computing
-the coefficients $c_{\v n k}$, there is only negligible computational
-overhead to compute the correlator $\bk{\S_{\v n}\p{t}}$ for any time
-$t$ up to the breakdown time $t_M$ of an order-$M$ TST expansion.
-
-\section{Short-time linear response and two-time correlators}
-\label{sec:linear_response}
-
-Here, we discuss the appearance of two-time correlation functions in
-the short-time linear response of correlators to perturbations of a
-Hamiltonian.  Consider an initial Hamiltonian $H$ perturbed by an
-operator $V$ with $\norm{V}\ll\norm{H}$, where $\norm{\O}$ denotes the
-operator norm of $\O$, such that the net Hamiltonian is
-$\widetilde H=H+V$.  We denote the generator of Heisenberg time
-evolution under the perturbed (unperturbed) Hamiltonian by
-$\widetilde T$ ($T$).  These generators are related by
-\begin{align}
-  \widetilde T = T + i\check V
-\end{align}
-where $\check V$ is a superoperator whose action on operators $\O$ is
-defined by
-\begin{align}
-  \check V \O \equiv \sp{V,\O}_-.
-\end{align}
-Through quadratic order in the time $t$ and linear order in the
-perturbation $\check V$, we can say that
-\begin{align}
-  e^{t\widetilde T}
-  \approx \f12\sp{e^{tT}, e^{it\check V}}_+
-  \approx e^{tT} + \f12 it \sp{e^{t T}, \check V}_+.
-\end{align}
-Defining perturbed and unperturbed Heisenberg operators
-$\widetilde\O\p{t}\equiv e^{t\widetilde T}\O$ and
-$\O\p{t}\equiv e^{tT}\O$, we thus find that for sufficiently small
-times $t$ and weak perturbations $V$,
-\begin{align}
-  \widetilde\O\p{t} - \O\p{t}
-  = \p{e^{t\widetilde T} - e^{tT}} \O
-  \approx \f12 i t \p{\sp{V,\O}_-\p{t} + \sp{V,\O\p{t}}_-}.
-  \label{eq:response}
-\end{align}
-Two-time operators $V\O\p{t}$ and $\O\p{t} V$, in addition to the
-Heisenberg operators $\p{V\O}\p{t}$ and $\p{\O V}\p{t}$, thus
-determine the short-time linear response of operators $\O\p{t}$ to
-perturbations $V$ of a Hamiltonian.
-
-% \section{Changing operator bases, and a general product for algebras
-%   of permutationally-symmetric operators}
-% \label{sec:new_basis}
-
-% For a collective spin system composed of many small spins, the
-% operators $\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$ contain
-% $\abs{\v m}$-body operators (for $\abs{\v m}\equiv\sum_\mu m_\mu$),
-% but also carry lots of ``baggage'' in the form of $k$-body operators
-% with $k<\abs{\v m}$.  To cut down on this overhead, we try to change
-% our basis for collective spin operators from $\S_{\v m}$ to the
-% purely-$\abs{\v m}$-body operators
-% \begin{align}
-%   \tilde\S_{\v m} \equiv \sum_{\p{\v j;\v m}} P_{\v j},
-%   &&
-%   P_{\v j} \equiv \prod_{\mu,a} s_\mu^{\p{j^\mu_a}},
-%   \label{eq:S_P}
-% \end{align}
-% where $\v m \equiv \p{m_+,m_\z,m_-}\in\mathbb{N}_0^3$ with $m_\mu$ the
-% number of $s_\mu$ operators in each term of $\S_{\v m}$; $\v j$ is an
-% ordered list of indices for the spins addressed by $P_{\v j}$; and the
-% sum over $\p{\v j;\v m}$ denotes a sum over all possible $\v j$ for
-% which all indices in $\v j$ are distinct and
-% $\abs{\set{j^\mu_a}} = m_\mu$, such that $j^\mu_a$ indexes the $a$-th
-% spin addressed by an $s_\mu$ operator in $P_{\v j}$.
-
-% In order to make use of the operators $\tilde S_{\v m}$, we will need
-% to compute products of the form $\tilde S_{\v m} \tilde S_{\v n}$.  We
-% therefore expand such a product into terms that have exactly $s$ spins
-% addressed by both operators:
-% \begin{align}
-%   \tilde\S_{\v m} \tilde\S_{\v n}
-%   = \sum_{s\ge0} \sum_{\p{\v j,\v k;\v m,\v n,s}} P_{\v j} P_{\v k},
-%   &&
-%   \sum_{\p{\v j,\v k;\v m,\v n,s}} X \equiv
-%   \sum_{\substack{\p{\v j;\v m},\p{\v k;\v n} \\
-%       \abs{\set{j_\alpha^\mu}\cap\set{k_\beta^\nu}}=s}} X.
-%   \label{eq:SS_PP}
-% \end{align}
-% Collecting terms in which $r_{\mu\nu}$ of the $s_\mu$ operators in
-% $P_{\v j}$ address the same spin as an $s_\nu$ operator in $P_{\v k}$,
-% we have
-% \begin{align}
-%   \tilde\S_{\v m} \tilde\S_{\v n}
-%   = \sum_{s\ge0} \sum_{\p{\v r;\v m,\v n,s}} f_{\v m\v n\v r}
-%   \sum_{\p{\v J,\v K;\v m,\v n,\v r}} P_{\v J} Q_{\v K},
-%   \label{eq:SS_PQ}
-% \end{align}
-% where the sum over $\p{\v r;\v m,\v n,s}$ denotes a sum over all
-% values of $\v r$ with the restrictions
-% \begin{align}
-%   r_{\mu\nu} \ge 0 ~\forall~ \mu,\nu,
-%   &&
-%   \sum_\nu r_{\mu\nu} \le m_\mu,
-%   &&
-%   \sum_\mu r_{\mu\nu} \le n_\nu,
-%   &&
-%   \sum_{\mu,\nu} r_{\mu\nu} = s;
-%   \label{eq:rest_r}
-% \end{align}
-% the factor $f_{\v m\v n\v r}$ counts the number of ways to pair
-% operators in $P_{\v j}$ and $P_{\v k}$ according to $\v r$, or
-% \begin{align}
-%   f_{\v m\v n\v r}
-%   &\equiv \sp{\prod_{\mu,\nu}
-%     { m_\mu - \sum_{\rho<\nu} r_{\mu\rho} \choose r_{\mu\nu} }
-%     { n_\mu - \sum_{\rho<\nu} r_{\rho\mu} \choose r_{\nu\mu} }}
-%   \sp{\prod_{\mu,\nu} r_{\mu\nu}!} \\
-%   &= \sp{\prod_\mu \f{m_\mu!}{\p{m_\mu-\sum_\rho r_{\mu\rho}}!}
-%     \f{n_\mu!}{\p{n_\mu-\sum_\rho r_{\rho\mu}}!}}
-%   \sp{\prod_{\mu,\nu} r_{\mu\nu}!}^{-1};
-% \end{align}
-% $\v J$ and $\v K$ are respectively ordered lists of indices for spins
-% addressed by one and two single-spin operators in a term of the
-% product $P_{\v j} P_{\v k}$ with fixed $\v r$; the sum over
-% $\p{\v J,\v K;\v m,\v n,\v r}$ denotes a sum over all values of
-% $\v J,\v K$ with the restrictions
-% \begin{align}
-%   \abs{\set{J_a^\mu}}
-%   = m_\mu + n_\mu - \sum_\rho\p{r_{\mu\rho}+r_{\rho\mu}},
-%   &&
-%   \abs{\set{K^{\mu\nu}_a}} = r_{\mu\nu},
-%   &&
-%   \set{J_a^\lambda} \cap \set{K_b^{\mu\nu}} = \varnothing,
-% \end{align}
-% with $\varnothing$ denoting the empty set; and finally, similarly to
-% $P_{\v j}$ in \eqref{eq:S_P} we define
-% \begin{align}
-%   Q_{\v K} \equiv \prod_{\mu,\nu,a}
-%   s_\mu^{(K^{\mu\nu}_a)} s_\nu^{(K^{\mu\nu}_a)}.
-% \end{align}
-% As the sum in \eqref{eq:SS_PQ} is invariant under permutation of the
-% indices in $\v K$, we can safely neglect keeping track of these
-% indices and simply write
-% \begin{align}
-%   Q_{\v K}
-%   = \prod_{\mu,\nu} \prod_{a=1}^{r_{\mu\nu}}
-%   s_\mu^{(K^{\mu\nu}_a)} s_\nu^{(K^{\mu\nu}_a)}
-%   \simeq \bigotimes_{\mu,\nu} \bigotimes_{a=1}^{r_{\mu\nu}} s_\mu s_\nu
-%   = \bigotimes_{\mu,\nu} \bigotimes_{a=1}^{r_{\mu\nu}}
-%   \sum_\rho \eta_{\mu\nu\rho} s_\rho,
-%   \label{eq:Q_K_eta}
-% \end{align}
-% where $\simeq$ denotes equality up to a re-indexing of spins and
-% additional tensor factors of the single-spin identity operator
-% (i.e.~$\1$); we have introduced explicit dependence on
-% $r_{\mu\nu}=\abs{\set{K^{\mu\nu}_a}}$ for brevity; and
-% $\eta_{\mu\nu\rho}$ is a structure constant defined by
-% $s_\mu s_\nu=\sum_\rho\eta_{\mu\nu\rho}s_\rho$.  Unlike the sums over
-% $\mu,\nu$ in most of the work above, the sum over $\rho$ here includes
-% an index for the identity operator $s_0\equiv\1$.  Distributing the
-% product of sums into a sum of products gives us
-% \begin{align}
-%   Q_{\v K}
-%   \simeq \sum_{\v\rho} \bigotimes_{\mu,\nu,a}
-%   \eta_{\mu\nu\rho^{\mu\nu}_a} s_{\rho^{\mu\nu}_a}
-%   = \sum_{\v\rho} \sp{\prod_{\mu,\nu,a}\eta_{\mu\nu\rho^{\mu\nu}_a}}
-%   \bigotimes_{\mu,\nu,a} s_{\rho^{\mu\nu}_a},
-%   \label{eq:Q_K_rho}
-% \end{align}
-% where implicitly $\abs{\set{\rho^{\mu\nu}_a}}=r_{\mu\nu}$, as
-% $\p{\mu,\nu,a}$ essentially index a factor in \eqref{eq:Q_K_eta} and
-% $\rho^{\mu\nu}_a$ indexes one of the terms in that factor.  Letting
-% $\tilde\rho^{\mu\nu}_\kappa$ denote the number of elements in the
-% index list
-% $\p{\rho^{\mu\nu}_1,\rho^{\mu\nu}_2,\cdots,\rho^{\mu\nu}_{r_{\mu\nu}}}$
-% that are equal to $\kappa$, we observe that two terms in
-% \eqref{eq:Q_K_rho} with, say, $\v\rho=\v\rho_1$ and $\v\rho=\v\rho_2$
-% are equal up to a permutation of indices if
-% $\tilde{\v\rho}_1=\tilde{\v\rho}_2$.  The degeneracy (under
-% permutation) of terms in \eqref{eq:Q_K_rho}, i.e.~the number of
-% $\v\rho$ that are consistent with $\tilde{\v\rho}$, is
-% \begin{align}
-%   g_{\tilde{\v\rho}}
-%   \equiv \prod_{\mu,\nu,\kappa}
-%   { r_{\mu\nu} - \sum_{\lambda<\kappa} \tilde\rho^{\mu\nu}_\lambda
-%     \choose \tilde\rho^{\mu\nu}_\kappa }
-%   = \sp{\prod_{\mu,\nu} r_{\mu\nu}!}
-%   \sp{\prod_{\mu,\nu,\kappa} \tilde\rho^{\mu\nu}_\kappa!}^{-1}.
-% \end{align}
-% Defining additionally
-% \begin{align}
-%   \v\eta^{\tilde{\v\rho}}
-%   \equiv \prod_{\mu,\nu,\kappa}
-%   \p{\eta_{\mu\nu\kappa}}^{\tilde\rho^{\mu\nu}_\kappa},
-%   \label{eq:eta_rho}
-% \end{align}
-% we can collect like factors and group together equivalent terms in
-% \eqref{eq:Q_K_rho} to write
-% \begin{align}
-%   Q_{\v K}
-%   \simeq \sum_{\tilde{\v\rho}} g_{\tilde{\v\rho}} \v\eta^{\tilde{\v\rho}}
-%   \bigotimes_{\kappa}
-%   s_\kappa^{\otimes\sum_{\mu,\nu}\tilde\rho^{\mu\nu}_\kappa}
-%   \simeq \sum_{\tilde{\v\rho}} g_{\tilde{\v\rho}} \v\eta^{\tilde{\v\rho}}
-%   \1^{\otimes\sum_{\mu,\nu}\tilde\rho^{\mu\nu}_0}
-%   \otimes \bigotimes_{\kappa\ne0}
-%   s_\kappa^{\otimes\sum_{\mu,\nu}\tilde\rho^{\mu\nu}_\kappa},
-%   \label{eq:Q_K_rho_simp}
-% \end{align}
-% where again implicitly
-% $\sum_\kappa\tilde\rho^{\mu\nu}_\kappa=r_{\mu\nu}$ for consistency
-% with \eqref{eq:Q_K_eta} and \eqref{eq:Q_K_rho}, and we have explicitly
-% factored out the identity operators in each term of $Q_{\v K}$.
-
-% We have now simplified $Q_{\v K}$ sufficiently to substitute it back
-% into \eqref{eq:SS_PQ}, which (removing the tilde from $\tilde{\v\rho}$
-% to simplify notation) gives us
-% \begin{align}
-%   \tilde\S_{\v m} \tilde\S_{\v n}
-%   = \sum_{s\ge0} \sum_{\p{\v r;\v m,\v n,s}} f_{\v m\v n\v r}
-%   \sum_{\p{\v\rho;\v r}} g_{\v\rho} \v\eta^{\v\rho}
-%   c_{\v\ell_{\v m\v n\v r\v\rho}\v\rho}
-%   \tilde\S_{\v\ell_{\v m\v n\v r\v\rho}},
-%   \label{eq:SS_simp}
-% \end{align}
-% where the sum over $\p{\v\rho;\v r}$ denotes a sum over all values of
-% $\v\rho$ and with the restrictions
-% \begin{align}
-%   \abs{\set{\rho^{\mu\nu}_\kappa}} = \abs{\set{s_\kappa}},
-%   &&
-%   \sum_{\kappa} \rho^{\mu\nu}_\kappa = r_{\mu\nu};
-% \end{align}
-% the combinatorial factor
-% \begin{align}
-%   c_{\v\ell\v\rho}
-%   \equiv \f{\p{N - \abs{\v\ell}}!}
-%   {\p{N - \abs{\v\ell} - \sum_{\mu,\nu}\rho^{\mu\nu}_0}!}
-% \end{align}
-% accounts for the number of terms that are equivalent up to a
-% permutation of indices involving the identity operators in
-% \eqref{eq:Q_K_rho_simp}; and the number of $s_\mu$ operators in each
-% term of $\tilde\S_{\v\ell_{\v m\v n\v r\v\rho}}$ is
-% \begin{align}
-%   \ell_{\v m\v n\v r\v\rho;\mu}
-%   = m_\mu + n_\mu - \sum_\kappa \p{r_{\mu\kappa}+r_{\kappa\mu}}
-%   + \sum_{\alpha,\beta} \rho^{\alpha\beta}_\mu.
-% \end{align}
-
-% The sums in \eqref{eq:SS_simp} generally involve a large number of
-% terms; in practice, many of these terms will be equal to zero.  We
-% can, however, impose additional restrictions on these sums in such a
-% way as to throw out all terms that are zero, and keep only those that
-% are not.  Remembering that $s$ counts the number of spins that are
-% addressed by both operators $P_{\v j}$ and $P_{\v k}$ in
-% \eqref{eq:SS_PP}, the first restriction we can impose comes from
-% recognizing that for given $\tilde\S_{\v m},\tilde\S_{\v n}$, the
-% overlap $s$ is bounded as
-% \begin{align}
-%   \max\set{0,\abs{\v m}+\abs{\v n}-N}
-%   \le s \le \min\set{N,\abs{\v m}+\abs{\v n}}.
-% \end{align}
-% In addition, the matrix of structure constants $\eta_{\mu\nu\kappa}$
-% may generally contain zeros that we can preemptively avoid with
-% restrictions on $\v r$ and $\v\rho$.  Recalling that $r_{\mu\nu}$
-% counts the number of operators addressed by
-% $s_\mu s_\nu=\sum_\kappa\eta_{\mu\nu\kappa} s_\kappa$, if
-% $\eta_{\mu\nu\kappa}=0$ for all $\kappa$, i.e.~$s_\mu s_\nu=0$, then
-% any term with $r_{\mu\nu}>0$ will vanish.  Our second restriction on
-% the sums in \eqref{eq:SS_simp} is therefore to fix $r_{\mu\nu}=0$ for
-% all $\mu,\nu$ with $s_\mu s_\nu=0$.  Finally, we can avoid the
-% nullifying effect of zeros from the factor
-% $\v\eta^{\v\rho}=\prod_{\mu,\nu,\kappa}
-% \p{\eta_{\mu\nu\kappa}}^{\rho^{\mu\nu}_\kappa}$ by fixing
-% $\rho^{\mu\nu}_\kappa=0$ for all $\mu,\nu,\kappa$ with
-% $\eta_{\mu\nu\kappa}=0$.  These restrictions are sufficient to ensure
-% that all individual terms in \eqref{eq:SS_simp} are nonzero, although
-% they do not rule out the possibility of several terms canceling out.
-
-% TODO: comment on the use of this new basis vs.~the old one.
-
-\bibliography{\jobname}
-
-\end{document}

### 2019-09-11 soc_squeezing/collective_paper/changing_bases.tex
added comments on cumulant expansion added more discussion of the breakdown of TST added appendix on Heisenberg operators in open quantum systems, and the 'noise' operators

--- 2018-12-04 soc_squeezing/paper/collective_spin.tex
@@
-\documentclass[pra,reprint,longbibliography]{revtex4-1}
-
-% linking references
-\usepackage{hyperref}
-\hypersetup{
-  breaklinks=true,
-  colorlinks=true,
-  linkcolor=blue,
-  filecolor=magenta,
-  urlcolor=cyan,
-}
-
-%%% symbols, notations, etc.
-\usepackage{physics,braket,bm,commath,amssymb}
-\renewcommand{\t}{\text} % text in math mode
-\newcommand{\f}[2]{\dfrac{#1}{#2}} % shorthand for fractions
-\newcommand{\p}[1]{\left(#1\right)} % parenthesis
-\renewcommand{\sp}[1]{\left[#1\right]} % square parenthesis
-\renewcommand{\set}[1]{\left\{#1\right\}} % curly parenthesis
-\renewcommand{\v}{\bm} % bold vectors
-\newcommand{\uv}[1]{\v{\hat{#1}}} % unit vectors
-\renewcommand{\c}{\cdot} % inner product
-\newcommand{\bk}{\Braket} % shorthand for braket notation
-
-\newcommand{\C}{\mathcal{C}}
-\newcommand{\D}{\mathcal{D}}
-\newcommand{\F}{\mathcal{F}}
-\newcommand{\I}{\mathcal{I}}
-\newcommand{\J}{\mathcal{J}}
-\renewcommand{\O}{\mathcal{O}}
-\renewcommand{\S}{\mathcal{S}}
-
-\newcommand{\N}{\mathbb{N}}
-
-\newcommand{\z}{\text{z}}
-\newcommand{\x}{\text{x}}
-\newcommand{\y}{\text{y}}
-\newcommand{\Z}{\text{Z}}
-\newcommand{\X}{\text{X}}
-\newcommand{\Y}{\text{Y}}
-\newcommand{\bmu}{{\bar\mu}}
-\newcommand{\bnu}{{\bar\nu}}
-
-\usepackage{dsfont} % for identity operator
-\newcommand{\1}{\mathds{1}}
-
-\newcommand{\up}{\uparrow}
-\newcommand{\dn}{\downarrow}
-
-\usepackage[inline]{enumitem} % for inline enumeration
-
-%%% figures
-\usepackage{graphicx} % for figures
-\graphicspath{{./figures/}} % set path for all figures
-
-% for strikeout text
-% normalem included to prevent underlining titles in the bibliography
-\usepackage[normalem]{ulem}
-
-% for leaving notes in the text
-\newcommand{\note}[1]{\textcolor{red}{#1}}
-
-\begin{document}
-
-\title{Short-time simulation of open collective quantum spin systems}
-
-\author{Michael A. Perlin}
-\email{mika.perlin@gmail.com}
-\author{Ana Maria Rey}
-\affiliation{JILA, National Institute of Standards and Technology and
-  University of Colorado, 440 UCB, Boulder, Colorado 80309, USA}
-\affiliation{Center for Theory of Quantum Matter, 440 UCB, Boulder,
-  Colorado 80309, USA}
-\affiliation{Department of Physics, University of Colorado, 390 UCB,
-  Boulder, Colorado 80309, USA}
-
-\begin{abstract}
-  Abstract goes here.
-\end{abstract}
-
-\maketitle
-
-\section{Introduction}
-
-Collective spin systems are a versatile resource in quantum science
-for a range of applications including quantum-enhanced metrology,
-quantum simulation, and quantum information.  Theoretical interest in
-such systems dates back to the mid-twentieth century with the
-introduction of the Lipkin-Meshkov-Glick (LMG) model as a toy model
-for testing many-body approximation methods in contemporary nuclear
-physics\cite{lipkin1965validity, meshkov1965validity,
-  glick1965validity}.  Since then, collective system systems have
-additionally featured in both theoretical and experimental studies of
-quantum criticality\cite{latorre2005entanglement,
-  alcalde2007functional, morrison2008dissipationdriven,
-  morrison2008collective, sarandy2009classical, wang2012quantum,
-  kessler2012dissipative, majd2014lmg}, non-equilibrium
-phenomena\cite{walls1978nonequilibrium, morrison2008dynamical,
-  bhattacherjee2014nonequilibrium, klinder2015dynamical,
-  maghrebi2016nonequilibrium, zhiqiang2017nonequilibrium,
-  lang2018concurrence, smale2018observation}, and precision
-metrology\cite{wineland1992spin, itano1993quantum,
-  kitagawa1993squeezed, ma2011quantum, chen2010heralded,
-  takano2009spin, appel2009mesoscopic, schleier-smith2010states,
-  chen2011conditional, riedel2010atomchipbased, gross2010nonlinear,
-  zhong2010simplified, agarwal1997atomic, lau2014proposal,
-  huang2015quantum, wineland1992spin, kitagawa1993squeezed,
-  schleier-smith2010squeezing, leroux2010implementation,
-  bohnet2014reduced, cox2016deterministic, hosten2016measurement,
-  hosten2016quantum, lewis-swan2018robust, norcia2018cavitymediated,
-  jones2009magnetic}.  With the development of advanced AMO (atomic,
-molecular, and optical) trapping, cooling, and control techniques,
-collective spin systems have been experimentally realized in atomic
-ensembles\cite{takano2009spin, appel2009mesoscopic, chen2010heralded,
-  schleier-smith2010states, chen2011conditional}, Bose-Einstein
-condensates\cite{riedel2010atomchipbased, gross2010nonlinear,
-  klinder2015dynamical}, ultracold Fermi
-gasses\cite{martin2013quantum, smale2018observation}, trapped
-ions\cite{bohnet2016quantum}, optical
-cavities\cite{leroux2010implementation, bohnet2014reduced,
-  cox2016deterministic, hosten2016measurement, hosten2016quantum,
-  zhiqiang2017nonequilibrium, lewis-swan2018robust,
-  norcia2018cavitymediated}, and molecular
-sensors\cite{jones2009magnetic}.
-
-One of the primary motivations for studying collective spin systems is
-their application to quantum-enhanced metrology.  Quantum projection
-noise limits the error $\Delta\phi$ in the measurement of a phase
-angle $\phi$ with $N$ independent spins to
-$\Delta\phi\sim1/\sqrt{N}$\cite{wineland1992spin, itano1993quantum,
-  ma2011quantum}.  Collective spin systems enable a means to break
-through this limit via the preparation of many-body entangled states
-such as NOON\cite{jones2009magnetic, chen2010heralded,
-  zhong2010simplified}, spin-cat\cite{agarwal1997atomic,
-  lau2014proposal, huang2015quantum}, and spin-squeezed
-states\cite{wineland1992spin, kitagawa1993squeezed, takano2009spin,
-  appel2009mesoscopic, schleier-smith2010states,
-  schleier-smith2010squeezing, leroux2010implementation,
-  riedel2010atomchipbased, gross2010nonlinear, chen2011conditional,
-  bohnet2014reduced, cox2016deterministic, hosten2016measurement,
-  hosten2016quantum, lewis-swan2018robust, norcia2018cavitymediated,
-  ma2011quantum} that allow for measurement errors
-$\Delta\phi\sim1/N^\epsilon$ with $\epsilon>1/2$.  Generating such
-states typically requires nonlinear dynamics that are realized using
-long-range collisional, photon-, or phonon-mediated interactions.
-Although a truly collective spin model requires uniform, all-to-all
-interactions, as long as measurements do not distinguish between
-constituent particles even non-uniform systems can be effectively
-described by a uniform model with renormalized
-parameters\cite{hu2015entangled}.
-
-In the absence of decoherence, permutation symmetry and total spin
-conservation divide the total Hilbert space of a collective spin
-system into superseletion sectors that grow only linearly with system
-size $N$, thereby admitting efficient classical simulation of its
-dynamics.  Decoherence will generally violate total spin conservation
-and require the use of density operators, increasing the dimension of
-accessible state space to $\sim N^3$\cite{hartmann2016generalized,
-  xu2013simulating} and allowing for exact simulations of
-$N\lesssim100$ particles.  If decoherence is sufficiently weak,
-dynamics can be numerically solvable for $N\lesssim10^5$ particles via
-``quantum trajectory'' Monte Carlo methods\cite{plenio1998quantumjump,
-  zhang2018montecarlo} (also known as ``quantum jump'' or ``Monte
-Carlo wavefunction'' methods) that can reproduce all expectation
-values of interest.  When decoherence is strong, however, or when the
-number of jump operators grows extensively with system size (e.g.~for
-single-spin decay), these Monte Carlo methods can take a prohibitively
-long time to converge, as simulations become dominated by incoherent
-jumps that generate large numbers of distinct trajectories over which
-to average.
-
-In this work, we develop a new method to simulate short-time dynamics
-of collective spin systems with permutationally-symmetric decoherence.
-This method is based on a short-time expansion of exact solutions to
-the equations of motion for Heisenberg operators.  Evaluating this
-expansion requires use of the structure factors of a collective spin
-operator algebra; the calculation of these structure factors is one of
-the main technical results of this work, which we hope will enable
-future studies to develop new analytical insights into the structure
-and dynamics of collective spin systems.
-
-[Paragraph or so about benchmarking and the discussion section?]
-
-\section{Theory}
-
-In this section we provide the basic theory for our method to compute
-expectation values of collective spin operators, deferring lengthy
-derivations to the Appendices.  We consider systems composed of $N$
-distinct spin-1/2 particles.  Defining individual spin-1/2 operators
-$s_{n=\x,\y,\z}\equiv\sigma_n/2$ and
-$s_\pm\equiv s_\x\pm is_\y=\sigma_\pm$ with Pauli operators
-$\sigma_n$, we denote an operator which acts with $s_n$ on the spin
-indexed by $j$ and trivially (i.e.~with the identity $\1$) on all
-other spins by $s_n^{(j)}$.  We then define the collective spin
-operators $S_n\equiv\sum_{j=1}^Ns_n^{(j)}$ for
-$n\in\set{\x,\y,\z,+,-}$.  Identifying the set $\set{\S_{\v m}}$ as a
-basis for all collective spin operators, with
-$\v m\equiv\p{m_+,m_\z,m_-}\in\N_0^3$ and
-$\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$, we can expand all
-collective spin Hamiltonians in the form
-\begin{align}
-  H = \sum_{\v m} h_{\v m} \S_{\v m},
-  \label{eq:H_general}
-\end{align}
-where Hermiticity requires that $h_{\v m}^*=h_{\v m^*}$ for
-$\v m^*\equiv\p{m_-,m_\z,m_+}$.  The time evolution of a particular
-Heisenberg operator $\S_{\v m}$ under a Hamiltonian of the form in
-Eq.~\eqref{eq:H_general} is then given by
-\begin{align}
-  \f{d}{dt}\S_{\v m}
-  &= i\sum_{\v n}h_{\v n}\sp{\S_{\v n},\S_{\v m}}_-
-  + \sum_\J \gamma_\J \D\p{\J;\S_{\v m}} \\
-  &= \sum_{\v n} T_{\v m\v n} \S_{\v n},
-  \label{eq:EOM}
-\end{align}
-where $\sp{X,Y}_\pm\equiv XY\pm YX$; $\J$ is a set of jump operators
-with a corresponding decoherence rate $\gamma_\J$;
-\begin{align}
-  \D\p{\J;\O}
-  \equiv \sum_{J\in\J} \p{J^\dag \O J - \f12\sp{J^\dag J,\O}_+}
-\end{align}
-is the Heisenberg picture dissipator, or Lindblad superoperator, for
-all jump operators $J\in\J$; and $T_{\v m\v n}$ is a matrix element of
-the time derivative operator.  Decoherence via uncorrelated decay of
-individual spins at a rate $\gamma_-$, for example, would be described
-by the set of jump operators
-$\J_-=\set{\sigma_-^{(j)}:j=1,2,\cdots,N}$ with
-$\gamma_{\J_-}=\gamma_-$.  The commutator in Eq.~\eqref{eq:EOM} can be
-determined from the structure factors of $f_{\v n\v m\v\ell}$ of the
-general product
-$\S_{\v n}\S_{\v m}=\sum_{\v\ell}f_{\v n\v m\v\ell}S_{\v\ell}$ worked
-out in Appendices \ref{sec:identities}--\ref{sec:prod_general}, while
-the effects of decoherence from jump operators (i.e.~elements of $\J$)
-of the form $\gamma^{(j)}=\sum_n\gamma_ns_n^{(j)}$ and
-$\Gamma=\sum_n\Gamma_nS_n$ are respectively worked out in Appendices
-\ref{sec:sandwich_single}--\ref{sec:decoherence_collective}.  These
-ingredients are sufficient to compute matrix elements $T_{\v m\v n}$
-of the time derivative operator $d/dt$ in Eq.~\eqref{eq:EOM} in most
-cases of experimental relevance.
-
-The time derivative operator $d/dt$ will generally couple spin
-operators $\S_{\v m}$ to spin operators $\S_{\v n}$ with higher
-``weight'', i.e.~with $\abs{\v n}>\abs{\v m}$, where
-$\abs{\v\ell}\equiv\sum_\lambda\ell_\lambda$.  The growth of operator
-weight signifies the growth of many-body correlations.  In practice,
-keeping track of this growth will eventually require more
-computational resources than are available, meaning we must somehow
-truncate our equations of motion.  The simplest truncation strategy
-would be to take
-\begin{align}
-  \f{d}{dt} \S_{\v m}
-  \to \sum_{w\p{\v n}<W} T_{\v m\v n} \S_{\v n}
-  \label{eq:weight_truncation}
-\end{align}
-for some weight measure $w$, e.g.~$w\p{\v n}=\abs{\v n}$, and cutoff
-$W$.  The truncation in Eq.~\eqref{eq:weight_truncation} closes the
-system of differential equations defined by Eq.~\eqref{eq:EOM}, and
-allow us to solve it using standard numerical methods.  Initial
-conditions for this system of differential equations, i.e.~expectation
-values of collective spin operators with respect to states that are
-simple to prepare experimentally, are provided in Appendix
-\ref{sec:initial_conditions}.
-
-The truncation strategy in Eq.~\eqref{eq:weight_truncation} has a few
-limitations:
-\begin{enumerate*}[label=(\roman*)]
-\item simulating a system of differential equations for a large number
-  of operators can be quite time-consuming,
-\item the weight measure $w$ may need to be chosen carefully, as the
-  optimal measure is generally system-dependent,
-\item simulation results can only be trusted up to the time at which
-  the initial value of operators $\S_{\v m}$ with weight
-  $w\p{\v m}\ge W$ have a non-negligible contribution to expectation
-  values of interest.\label{pt:limitation}
-\end{enumerate*}
-The last limitation in particular unavoidably applies in some form to
-any method tracking only a subset of all relevant operators.  We
-therefore devise an alternate truncation strategy built around
-limitation \ref{pt:limitation}.
-
-We can expand Heisenberg operators in a Taylor series about $t=0$ to
-write
-\begin{align}
-  \bk{\S_{\v m}}
-  = \sum_{k\ge0} \f{t^k}{k!} \bk{\f{d^k}{dt^k} \S_{\v m}}_{t=0}
-  = \sum_{k\ge0} \f{t^k}{k!}
-  \sum_{\v n} T^k_{\v m\v n} \bk{\S_{\v n}}_{t=0},
-  \label{eq:time_series}
-\end{align}
-where the matrix elements $T^k_{\v m\v n}$ of the $k$-th time
-derivative are
-\begin{align}
-  T^0_{\v m\v n} &\equiv \delta_{\v m\v n}, \\
-  T^1_{\v m\v n} &\equiv T_{\v m\v n}, \\
-  T^{k>1}_{\v m\v n}
-  &\equiv \sum_{\v p_1,\v p_2,\cdots,\v p_{k-1}}
-  T_{\v m\v p_1} T_{\v p_1\v p_2} T_{\v p_2\v p_3}
-  \cdots T_{\v p_{k-1}\v n},
-\end{align}
-with $\delta_{\v m\v n}=1$ if $\v m=\v n$ and 0 otherwise.  For
-sufficiently short times, we can truncate the series in
-Eq.~\eqref{eq:time_series} by taking
-\begin{align}
-  \bk{\S_{\v m}}
-  \to \sum_{k=0}^M \f{t^k}{k!}
-  \sum_{\v n} T^k_{\v m\v n} \bk{\S_{\v n}}_{t=0}.
-  \label{eq:order_truncation}
-\end{align}
-Unlike the truncation in Eq.~\eqref{eq:weight_truncation}, the nonzero
-matrix elements $T^k_{\v m\v n}$ for $k=0,1,\cdots,M$ now tell us
-which operators are relevant for computing the expectation value
-$\bk{\S_{\v m}}$ to a fixed order $M$.  Note that the relation
-$\S_{\v m^*}=\S_{\v m}^\dag$, which by Hermitian conjugation of
-Eq.~\eqref{eq:EOM} also implies that
-$T_{\v m^*\v n^*}=T_{\v m\v n}^*$, can be exploited to cut both the
-number of expectation values $\bk{\S_{\v m}}_{t=0}$ and matrix
-elements $T_{\v m\v n}$ that we need to explicitly compute roughly in
-half.
-
-In principle, the factorial suppression of terms at higher orders
-(i.e.~in $k$) of the expansion in Eq.~\eqref{eq:order_truncation}
-implies for any given time $t$, there exists a truncation order $M_t$
-for which truncation error is negligibly small.  In practice, only a
-fixed maximal truncation order $\overline M$ is accessible with
-limited computational resources, such that
-Eq.~\eqref{eq:order_truncation} only allows computation of the
-correlator $\bk{\S_{\v m}}$ to a maximal time $t_{\overline M}$.  The
-maximal accessible truncation order $\overline M$ generally depends on
-the choice of operator basis for collective spin systems, which we
-discuss briefly at the end of Appendix \ref{sec:new_basis}.  As we
-will see in the following sections, our choice of basis elements
-$\S_{\v m}=S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$ allows us to accurately
-compute spin correlators to times that are relevant e.g.~for spin
-squeezing protocols.
-
-\section{Spin squeezing}
-
-To benchmark our method for computing collective spin correlators, we
-consider three collective spin models known to dynamically generate
-spin-squeezed states: the one-axis twisting (OAT), two-axis twisting
-(TAT), and twist-and-turn (TNT) models described by the collective
-spin Hamiltonians\cite{kitagawa1993squeezed, micheli2003manyparticle}
-\begin{align}
-  \label{eq:OAT}
-  H_{\t{OAT}} &= \chi S_\z^2, \\
-  \label{eq:TAT}
-  H_{\t{TAT}} &= \f{\chi}{3} \p{S_\z^2 - S_\y^2}, \\
-  \label{eq:TNT}
-  H_{\t{TNT}} &= \chi S_\z^2 + \Omega S_\x,
-\end{align}
-where we include a factor of $1/3$ in the TAT Hamiltonian because it
-naturally appears in realistic proposals to experimentally realize
-TAT\cite{liu2011spin, huang2015twoaxis}.  For simplicity, we further
-fix $\Omega=S\chi$ (with $S\equiv N/2$ throughout this work) to the
-critical value known to maximize the entanglement generation rate of
-of TNT in the large-$N$ limit\cite{micheli2003manyparticle,
-  sorelli2015fast}.  Note that the OAT model is a special case of the
-zero-field Ising model, whose quantum dynamics in the presence of
-decoherence via uncorrelated decay, excitation, and dephasing of
-individual spins have been previously solved
-analytically\cite{foss-feig2013nonequilibrium} (see Appendix
-\ref{sec:OAT}, as well as the Supplementary Material of
-Ref.~[\citenum{bohnet2016quantum}] for more general multi-spin
-correlators).  We will benchmark against these analytical OAT results
-wherever applicable.
-
-The Hamiltonians in Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT} squeeze the
-initial product state $\ket{\X}\propto\p{\ket\up+\ket\dn}^{\otimes N}$
-with $S_\x\ket{\X}=S\ket{\X}$.  Our measure of spin squeezing is the
-directionally-unbiased squeezing parameter determined by the maximal
-gain in resolution $\Delta\phi$ of a phase angle $\phi$ over that
-achieved by any spin-polarized product state
-(e.g.~$\ket{\X}$)\cite{wineland1992spin, ma2011quantum},
-\begin{align}
-  \xi^2
-  \equiv \f{\p{\Delta\phi_{\t{min}}}^2}{\p{\Delta\phi_{\t{polarized}}}^2}
-  = \f{N}{\abs{\bk{\v S}}^2}
-  \min_{\uv n\perp\bk{\v S}} \bk{\p{\v S\c\uv n}^2},
-  \label{eq:squeezing}
-\end{align}
-where $\v S\equiv\p{S_\z,S_\x,S_\y}$ is a collective spin vector, and
-the minimization is performed over all unit vectors $\uv n$ in the
-plane orthogonal to the mean spin vector $\bk{\v S}$.  This squeezing
-parameter is entirely determined by one- and two-spin correlators of
-the form $\bk{S_m}$ and $\bk{S_m S_n}$.  In the case of coherent
-dynamics under TAT or TNT, these correlators are obtainable via exact
-simulations of quantum dynamics in the $\p{N+1}$-dimensional Dicke
-manifold of states $\set{\ket{S,m}}$ with net spin $S$ and spin
-projection $m$ onto the $z$ axis, i.e.~with
-$\bk{S,m|\v S^2|S,m}=S\p{S+1}$ and $\bk{S,m|S_\z|S,m}=m$ for
-$\abs{m}\le S$.  In the presence of permutationally-symmetric
-decoherence, these correlators are obtainable with the collective-spin
-quantum trajectory Monte Carlo method recently developed in
-ref.~\cite{zhang2018montecarlo}.  These exact and quantum trajectory
-simulations will serve as our methods for benchmarking the expansion
-in Eq.~\eqref{eq:order_truncation}.
-
-\begin{figure}
-  \centering
-  \includegraphics{coherent.pdf}
-  \caption{Spin squeezing of $N=10^4$ spins initially in $\ket\X$
-    under coherent evolution, as computed using exact methods (solid
-    lines) and the expansion in Eq.~\eqref{eq:order_truncation} with
-    $M=40$ (dashed lines).  Solid circles mark the times for TAT and
-    TNT after which the latter method gives an unphysical result with
-    $\xi^2<0$.  Both methods agree exactly in the OAT case for all
-    times shown.}
-  \label{fig:coherent}
-\end{figure}
-
-\begin{figure}
-  \centering
-  % \includegraphics{weak_decoherence.pdf}
-  \caption{Spin squeezing of $N=10^4$ spins initially in $\ket\X$ with
-    decoherence via uncorrelated decay, excitation, and dephasing of
-    individual spins at rates $\gamma_-=\gamma_+=\gamma_\z=\chi$.
-    Squeezing is computed using an exact simulation (solid line),
-    quantum trajectory methods (dots), and the expansion in
-    Eq.~\eqref{eq:order_truncation} with $M=40$ (dashed lines).  Solid
-    circles mark the times for TAT and TNT after which the latter
-    method gives an unphysical result with $\xi^2<0$.}
-  \label{fig:weak_decoherence}
-\end{figure}
-
-Figure \ref{fig:coherent} compares the squeezing parameter $\xi^2$ for
-$N=10^4$ spins initially in the state $\ket{\X}$ coherently evolved
-under the Hamiltonians in Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT}, as
-computed via both benchmarking simulations and the prescription in
-Eq.~\eqref{eq:order_truncation} with $M=40$.  Similarly, Figure
-\ref{fig:weak_decoherence} compares squeezing in the same scenario
-with the addition of decoherence through uncorrelated emission,
-absorption, and dephasing of individual spins at a rate $\chi$,
-respectively described by the sets of jump operators
-$\J_n=\set{s_n^{(j)}}$ with corresponding decoherence rates
-$\gamma_n=\chi$ for $n\in\set{-,+,\z}$.  The squeezing results shown
-in Figures \ref{fig:coherent} and \ref{fig:weak_decoherence} were
-computed in a rotated basis with $\p{s_\z,s_\x}\to\p{s_\x,-s_\z}$ and
-$\ket\X\to\ket{-\Z}\equiv\ket\dn^{\otimes N}$; the only effect of this
-rotation on the results presented in these figures is to roughly
-double the time span of near-perfect agreement in squeezing via TNT.
-The reason for different results in rotated bases has to do with the
-breakdown of the expansion in Eq.~\eqref{eq:order_truncation}, which
-we discuss further in Section \ref{sec:breakdown}.
-
-[Comment on near-perfect agreement right up until a sudden and drastic
-departure from the correct result.  Reference figure showing squeezing
-for large system with strong decoherence.  Mention that this method
-was used to put a lower bound on squeezing via TAT in our
-SOC+interactions squeezing paper.]
-
-\section{Many-body correlators}
-
-Many-body correlators and cumulants \`a la
-Ref.~[\citenum{schweigler2017solving}]?
-
-\section{Breakdown}
-\label{sec:breakdown}
-
-\section{Conclusions}
-
-\newpage
-\onecolumngrid
-\appendix
-
-\section{Exact results for the one-axis twisting model}
-\label{sec:OAT}
-
-The one-axis twisting (OAT) Hamiltonian for $N$ spin-1/2 particles
-takes the form
-\begin{align}
-  H_{\t{OAT}}
-  = \chi S_\z^2
-  = \f12 \chi \sum_{j<k} \sigma_\z^{(j)} \sigma_\z^{(k)} + \f14 N \chi,
-\end{align}
-where $\sigma_\z^{(j)}$ represents a Pauli-$z$ operator acting on spin
-$j$.  This model is a special case of the zero-field Ising Hamiltonian
-previously solved in Ref.~[\citenum{foss-feig2013nonequilibrium}] via
-exact, analytical treatment of the quantum trajectory Monte Carlo
-method for computing expectation values.  The solution therein
-accounts for coherent evolution in addition to decoherence via
-uncorrelated single-spin decay, excitation, and dephasing respectively
-at rates $\gamma_-$, $\gamma_+$, and $\gamma_\z$ (denoted by
-$\Gamma_{\t{ud}}$, $\Gamma_{\t{du}}$, and $\Gamma_{\t{el}}$ in
-Ref.~[\citenum{foss-feig2013nonequilibrium}]).  Letting $S\equiv N/2$
-and $\mu,\nu\in\set{+1,-1}$, we adapt expectation values computed in
-Ref.~[\citenum{foss-feig2013nonequilibrium}] for the initial state
-$\ket{\X}$ satisfying $S_\x\ket{\X}=S\ket{\X}$ evolving under
-$H_{\t{OAT}}$, finding
-\begin{align}
-  \bk{S_+}
-  &= S e^{-\Gamma t} \Phi\p{\chi,t}^{N-1}, \label{eq:S+_OAT} \\
-  \bk{S_\mu S_\z}
-  &= -\f{\mu}{2}\bk{S_\mu} + S \p{S-\f12} e^{-\Gamma t}
-  \Psi\p{\mu\chi,t} \Phi\p{\chi,t}^{N-2}, \\
-  \bk{S_\mu S_\nu}
-  &= \delta_{\mu,-\nu} \p{S + \mu\bk{S_\z}}
-  + S \p{S-\f12} e^{-2\Gamma t}
-  \Phi\p{\sp{\mu+\nu}\chi,t}^{N-2}, \label{eq:SS+-_OAT}
-\end{align}
-where
-\begin{align}
-  \Phi\p{X,t}
-  &\equiv e^{-\lambda t} \sp{\cos\p{t\sqrt{s_X^2-r}}
-    + \lambda t~\t{sinc}\p{t\sqrt{s_X^2-r}}},
-  \\
-  \Psi\p{X,t}
-  &\equiv e^{-\lambda t} \p{is_X-\gamma}t~
-  \t{sinc}\p{t\sqrt{s_X^2-r}},
-\end{align}
-for
-\begin{align}
-  \gamma \equiv -\f12 \p{\gamma_+ - \gamma_-},
-  &&
-  \lambda \equiv \f12 \p{\gamma_+ + \gamma_-},
-  &&
-  r \equiv \gamma_+ \gamma_-,
-  &&
-  \Gamma \equiv \f12\gamma_\z + \lambda,
-  &&
-  s_X \equiv X + i\gamma,
-\end{align}
-In order to compute spin squeezing as measured by $\xi^2$ in
-\eqref{eq:squeezing}, we additionally need analytical expressions for
-$\bk{S_\z}$ and $\bk{S_\z^2}$.  As these operators commute with both
-the OAT Hamiltonian and the single-spin operators $\sigma_\z^{(j)}$,
-their evolution is governed entirely by decay-type decoherence (see
-Appendix \ref{sec:decay_single}), which means
-\begin{align}
-  \f{d}{dt} S_\z
-  &= S\p{\gamma_+-\gamma_-} - \p{\gamma_++\gamma_-} S_\z,
-  \\
-  \f{d}{dt}\p{S_\z^2}
-  &= S\p{\gamma_++\gamma_-} + 2\p{\gamma_+-\gamma_-}\p{S-\f12} S_\z
-  - 2 \p{\gamma_++\gamma_-} S_\z^2.
-\end{align}
-The initial conditions $\bk{S_\z}_{t=0}=0$ and $\bk{S_\z^2}_{t=0}=S/2$
-then imply
-\begin{align}
-  \bk{S_\z}
-  &= S\p{\f{\gamma_+-\gamma_-}{\gamma_++\gamma_-}}
-  \p{1-e^{-\p{\gamma_+ + \gamma_-} t}}, \\
-  \bk{S_\z^2} &= S \sp{\f12 + \p{S-\f12} \f{\bk{S_\z}^2}{S^2}}.
-  \label{eq:Sz_OAT}
-\end{align}
-The expectation values in \eqref{eq:S+_OAT}-\eqref{eq:SS+-_OAT} and
-\eqref{eq:Sz_OAT} are sufficient to compute the spin squeezing
-parameter $\xi^2$ in \eqref{eq:squeezing} at any time throughout
-evolution of the initial state $\ket{\X}$ under $H_{\t{OAT}}$, under
-appropriate assumptions about the relevant sources of decoherence.
-
-\section{Basic spin operator identities}
-\label{sec:identities}
-
-The appendices in this work make ubiquitous use of various spin
-operator identities; we collect and derive some basic identities here
-for reference.  Note that despite the working definition of collective
-spin operators from $S_n=\sum_js_n^{(j)}$, the identities we will
-derive involving only collective spin operators apply just as well to
-large-spin operators that cannot be expressed as the sum of individual
-spin-1/2 operators.  The elementary commutation relations between spin
-operators are, with $\bmu\equiv-\mu\in\set{+1,-1}$ for brevity,
-\begin{align}
-  \sp{s_\z^{(j)},s_\mu^{(k)}}_-
-  &= \delta_{jk} \mu s_\mu^{(j)},
-  &
-  \sp{S_\z,s_\mu^{(j)}}_-
-  &= \sp{s_\z^{(j)},S_\mu}_- = \mu s_\mu^{(j)},
-  &
-  \sp{S_\z,S_\mu}_-
-  &= \mu S_\mu,
-  \label{eq:comm_z_base} \\
-  \sp{s_\mu^{(j)},s_\bmu^{(k)}}_-
-  &= \delta_{jk} 2 \mu s_\z^{(j)},
-  &
-  \sp{S_\mu,s_\bmu^{(j)}}_-
-  &= \sp{s_\mu^{(j)},S_\bmu}_- = 2 \mu s_\z^{(j)},
-  &
-  \sp{S_\mu,S_\bmu}_-
-  &= 2 \mu S_\z.
-  \label{eq:comm_mu_base}
-\end{align}
-These relations can be used to inductively compute identities
-involving powers of collective spin operators.  By pushing through one
-spin operator at a time, we can find
-\begin{align}
-  \p{\mu S_\z}^m s_\mu^{(j)}
-  = \p{\mu S_\z}^{m-1} s_\mu^{(j)} \p{1 + \mu S_\z}
-  = \p{\mu S_\z}^{m-2} s_\mu^{(j)} \p{1 + \mu S_\z}^2
-  = \cdots
-  = s_\mu^{(j)} \p{1 + \mu S_\z}^m,
-  \label{eq:push_z_mu_Ss}
-\end{align}
-and
-\begin{align}
-  \mu s_\z^{(j)} S_\mu^m
-  = S_\mu \mu s_\z^{(j)} S_\mu^{m-1} + s_\mu^{(j)} S_\mu^{m-1}
-  = \cdots
-  = S_\mu^m \mu s_\z^{(j)} + ms_\mu^{(j)} S_\mu^{m-1},
-  \label{eq:push_z_mu_sS}
-\end{align}
-where we will generally find it nicer to express results in terms of
-$\mu s_\z^{(j)}$ and $\mu S_\z$ rather than $s_\z^{(j)}$ and $S_\z$.
-Summing over the single-spin index $j$ in both of the cases above
-gives us the purely collective-spin versions of these identities:
-\begin{align}
-  \p{\mu S_\z}^m S_\mu = S_\mu \p{1 + \mu S_\z}^m,
-  &&
-  \mu S_\z S_\mu^m = S_\mu^m \p{m + \mu S_\z},
-  \label{eq:push_z_mu_single}
-\end{align}
-where we can repeat the process of pushing through individual $S_\z$
-operators $\ell$ times to get
-\begin{align}
-  \p{\mu S_\z}^\ell S_\mu^m
-  = \p{\mu S_\z}^{\ell-1} S_\mu^m \p{m + \mu S_\z}
-  = \p{\mu S_\z}^{\ell-2} S_\mu^m \p{m + \mu S_\z}^2
-  = \cdots
-  = S_\mu^m \p{m + \mu S_\z}^\ell.
-  \label{eq:push_z_mu}
-\end{align}
-Multiplying \eqref{eq:push_z_mu} through by $\p{\mu\nu}^\ell$ (for
-$\nu\in\set{+1,-1}$) and taking its Hermitian conjugate, we can say
-that more generally
-\begin{align}
-  \p{\nu S_\z}^\ell S_\mu^m
-  = S_\mu^m \p{\mu\nu m+\nu S_\z}^\ell,
-  &&
-  S_\mu^m \p{\nu S_\z}^\ell
-  = \p{-\mu\nu m+\nu S_\z}^\ell S_\mu^m.
-\end{align}
-Finding commutation relations between powers of transverse spin
-operators, i.e.~$S_\mu$ and $S_\bmu$, turns out to be considerably
-more difficult than the cases we have worked out thus far.  We
-therefore save this work for Appendix \ref{sec:comm_transverse}.
-
-\section{Commutation relations between powers of transverse spin
-  operators}
-\label{sec:comm_transverse}
-
-To find commutation relations between powers of transverse collective
-spin operators, we first compute
-\begin{align}
-  S_\mu^m s_\bmu^{(j)}
-  &= S_\mu^{m-1}s_\bmu^{(j)} S_\mu
-  + S_\mu^{m-1} 2\mu s_\z^{(j)} \\
-  &= S_\mu^{m-2} s_\bmu^{(j)} S_\mu^2
-  + S_\mu^{m-2} 2\mu s_\z^{(j)} S_\mu
-  + S_\mu^{m-1} 2\mu s_\z^{(j)} \\
-  &= s_\bmu^{(j)} S_\mu^m
-  + \sum_{k=0}^{m-1} S_\mu^k 2\mu s_\z^{(j)} S_\mu^{m-k-1}
-  \label{eq:push_mu_Ss_start}.
-\end{align}
-While \eqref{eq:push_mu_Ss_start} gives us the commutator
-$\sp{S_\mu^m,s_\bmu^{(j)}}_-$, we would like to enforce an ordering on
-products of spin operators, which will ensure that we only keep track
-of operators that are linearly independent.  We choose (for now) to
-impose an ordering with all $s_\bmu^{(j)}$ operators on the left, and
-all $s_\z^{(j)}$ operators on the right, as such an ordering will
-prove convenient for the calculations in this section\footnote{In
-  retrospect, it would likely have been nicer to have all
-  $s_\mu^{(j)}$ operators on the right throughout these calculations,
-  due to the enhanced symmetry of expressions with respect to
-  Hermitian conjugation.  In any case, we express the final result of
-  this section in both ordering conventions, and thus feel no need to
-  reproduce these calculations with a different ordering.}.  This
-choice of ordering compels us to expand
-\begin{align}
-  \sum_{k=0}^{m-1} S_\mu^k 2\mu s_\z^{(j)} S_\mu^{m-k-1}
-  &= \sum_{k=0}^{m-1} S_\mu^k
-  \sp{2\p{m-k-1} s_\mu^{(j)} S_\mu^{m-k-2}
-    + S_\mu^{m-k-1} 2\mu s_\z^{(j)}} \\
-  &= m \p{m-1} s_\mu^{(j)} S_\mu^{m-2}
-  + m S_\mu^{m-1} 2\mu s_\z^{(j)},
-\end{align}
-which implies
-\begin{align}
-  S_\mu^m s_\bmu^{(j)}
-  = s_\bmu^{(j)} S_\mu^m + m \p{m-1} s_\mu^{(j)} S_\mu^{m-2}
-  + m S_\mu^{m-1} 2\mu s_\z^{(j)},
-  \label{eq:push_mu_Ss}
-\end{align}
-and in turn
-\begin{align}
-  S_\mu^m S_\bmu = S_\bmu S_\mu^m
-  + m S_\mu^{m-1} \p{m - 1 + 2\mu S_\z}.
-  \label{eq:push_mu_single}
-\end{align}
-As the next logical step, we take on the task of computing
-\begin{align}
-  S_\mu^m S_\bmu^n
-  = S_\mu^{m-1} S_\bmu^n S_\mu
-  + n \sp{S_\mu^{m-1} S_\bmu^{n-1} \p{1 - n + 2\mu S_\z}}
-  = S_\bmu^n S_\mu^m
-  + n \sum_{k=0}^{m-1} S_\mu^{m-k-1} S_\bmu^{n-1}
-  \p{1 - n + 2\mu S_\z} S_\mu^k,
-\end{align}
-which implies
-\begin{align}
-  \sp{S_\mu^m, S_\bmu^n}_-
-  = C_{mn;\mu}
-  \equiv n \sum_{k=0}^{m-1} S_\mu^{m-k-1} S_\bmu^{n-1}
-  \p{1 - n + 2\mu S_\z} S_\mu^k.
-\end{align}
-We now need rearrange the operators in $C_{mn;\mu}$ into a standard
-order, which means pushing all $S_\z$ operators to the right and, for
-the purposes of this calculation, all $S_\bmu$ operators to the left.
-We begin by pushing $S_\mu^k$ to the left of $S_\z$, which takes
-$2\mu S_\z\to 2\mu S_\z+2k$, and then push $S_\mu^{m-k-1}$ to the
-right of $S_\bmu^{n-1}$, giving us
-\begin{align}
-  C_{mn;\mu}
-  &= n \sum_{k=0}^{m-1}
-  \p{S_\bmu^{n-1} S_\mu^{m-k-1} + C_{m-k-1,n-1;\mu}} S_\mu^k
-  \p{2k + 1 - n + 2\mu S_\z} \\
-  &= D_{mn;\mu}
-  + n \sum_{k=0}^{m-2} C_{m-k-1,n-1;\mu}
-  S_\mu^k \p{2k + 1 - n + 2\mu S_\z},
-  \label{eq:C_mn}
-\end{align}
-where we have dropped the last ($k=m-1$) term in the remaining sum
-because $C_{m-k-1,n-1;\mu}=0$ if $k=m-1$, and
-\begin{align}
-  D_{mn;\mu}
-  \equiv mn S_\bmu^{n-1} S_\mu^{m-1} \p{m - n + 2\mu S_\z}.
-  \label{eq:D_mn}
-\end{align}
-To our despair, we have arrived in \eqref{eq:C_mn} at a {\it
-  recursive} formula for $C_{mn;\mu}$.  Furthermore, we have not even
-managed to order all spin operators, as $C_{m-k-1,n-1;\mu}$ contains
-$S_\z$ operators which are to the left of $S_\mu^k$.  To sort all spin
-operators once and for all, we define
-\begin{align}
-  C_{mn;\mu}^{(k)} \equiv C_{m-k,n;\mu} S_\mu^k,
-  &&
-  D_{mn;\mu}^{(k)} \equiv D_{m-k,n;\mu} S_\mu^k,
-\end{align}
-which we can expand as
-\begin{align}
-  D_{mn;\mu}^{(k)}
-  &= \p{m-k}n S_\bmu^{n-1} S_\mu^{m-k-1}
-  \p{m-k-n+2\mu S_\z} S_\mu^k \\
-  &= \p{m-k}n S_\bmu^{n-1} S_\mu^{m-1} \p{k+m-n+2\mu S_\z},
-  \label{eq:D_mn_k}
-\end{align}
-and
-\begin{align}
-  C_{mn;\mu}^{(k)}
-  &= D_{m-k,n;\mu} S_\mu^k + n \sum_{j=0}^{m-k-2}
-  C_{m-k-j-1,n-1;\mu} S_\mu^j \p{2j+1-n+2\mu S_\z} S_\mu^k \\
-  &= D_{mn;\mu}^{(k)} + n \sum_{j=0}^{m-k-2}
-  C_{m-k-j-1,n-1;\mu} S_\mu^{j+k} \p{2j+2k+1-n+2\mu S_\z} \\
-  &= D_{mn;\mu}^{(k)} + n \sum_{j=0}^{m-k-2}
-  C_{m-1,n-1;\mu}^{(k+j)} \p{2\sp{j+k}+1-n+2\mu S_\z} \\
-  &= D_{mn;\mu}^{(k)} + n \sum_{j=k}^{m-2}
-  C_{m-1,n-1;\mu}^{(j)} \p{2j+1-n+2\mu S_\z}.
-  \label{eq:C_mn_k}
-\end{align}
-While the resulting expression in \eqref{eq:C_mn_k} strongly resembles
-that in \eqref{eq:C_mn}, there is one crucial difference: all spin
-operators in \eqref{eq:C_mn_k} have been sorted into a standard order.
-We can now repeatedly substitute $C_{mn;\mu}^{(k)}$ into itself, each
-time decreasing $m$ and $n$ by 1, until one of $m$ or $n$ reaches
-zero.  Such repeated substitution yields the expansion
-\begin{align}
-  C_{mn;\mu}
-  = C_{mn;\mu}^{(0)}
-  = D_{mn;\mu}
-  + \sum_{p=1}^{\min\set{m,n}-1} E_{mn;\mu}^{(p)},
-  \label{eq:C_mn_E}
-\end{align}
-where the first two terms in the sum over $p$ are
-\begin{align}
-  E_{mn;\mu}^{(1)}
-  &= n \sum_{k=0}^{m-2} D_{m-1,n-1;\mu}^{(k)} \p{2k+1-n+2\mu S_\z}, \\
-  E_{mn;\mu}^{(2)}
-  &= n \sum_{k_1=0}^{m-2} \p{n-1} \sum_{k_2=k_1}^{m-3}
-  D_{m-2,n-2;\mu}^{(k_2)} \p{2k_2+2-n+2\mu S_\z} \p{2k_1+1-n+2\mu S_\z},
-\end{align}
-and more generally for $p>1$,
-\begin{align}
-  E_{mn;\mu}^{(p)}
-  = \f{n!}{\p{n-p}!}
-  \sum_{k_1=0}^{m-2} \sum_{k_2=k_1}^{m-3} \cdots\sum_{k_p=k_{p-1}}^{m-p-1}
-  D_{m-p,n-p;\mu}^{(k_p)} \prod_{j=1}^p \p{2k_j+j-n+2\mu S_\z}.
-  \label{eq:E_mn_p}
-\end{align}
-In principle, the expressions in \eqref{eq:D_mn_k} and
-\eqref{eq:C_mn_E}-\eqref{eq:E_mn_p} can be used to evaluate the
-commutator $\sp{S_\mu^m,S_\bmu^n}_- = C_{mn;\mu}$, but these result
-are -- to put it lightly -- quite a mess: the expression for
-$E_{mn;\mu}^{(p)}$ in \eqref{eq:E_mn_p} involves a sum over $p$
-mutually dependent intermediate variables, each term of which
-additionally contains a product of $p$ factors.  We therefore devote
-the rest of this section to simplifying our result for the commutator
-$\sp{S_\mu^m,S_\bmu^n}_-$.
-
-Observing that in \eqref{eq:E_mn_p} we always have
-$0\le k_1\le k_2\le\cdots\le k_p\le m-p-1$, we can rearrange the order
-of the sums and relabel $k_p\to\ell$ to get
-\begin{align}
-  E_{mn;\mu}^{(p)}
-  = \f{n!}{\p{n-p}!}
-  \sum_{\ell=0}^{m-p-1} D_{m-p,n-p;\mu}^{(\ell)} \p{2\ell+F_{np;\mu}}
-  \sum_{\p{\v k,p-1,\ell}} \prod_{j=1}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}},
-  \label{eq:E_mn_p_sum}
-\end{align}
-where for shorthand we define
-\begin{align}
-  F_{np;\mu} \equiv p - n + 2\mu S_\z,
-  &&
-  \sum_{\p{\v k,q,\ell}} X \equiv
-  \sum_{k_1=0}^\ell \sum_{k_2=k_1}^\ell
-  \cdots \sum_{k_q=k_{q-1}}^\ell X.
-\end{align}
-We now further define
-\begin{align}
-  f_{np\ell;\mu}\p{k,q} \equiv \p{\ell-k+q} \p{\ell+k-q+F_{np;\mu}},
-\end{align}
-and evaluate sums successively over $k_{p-1},k_{p-2},\cdots,k_1$,
-finding
-\begin{align}
-  \sum_{\p{\v k,p-1,\ell}} \prod_{j=1}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}}
-  &= \sum_{\p{\v k,p-2,\ell}}
-  \prod_{j=2}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}}
-  f_{np\ell;\mu}\p{k_{p-2},1} \\
-  &= \f1{\p{r-1}!} \sum_{\p{\v k,p-r,\ell}}
-  \prod_{j=r}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}}
-  \prod_{q=1}^{r-1} f_{np\ell;\mu}\p{k_{p-r},q} \\
-  &= \f1{\p{p-1}!} \prod_{q=1}^{p-1} f_{np\ell;\mu}\p{0,q} \\
-  &= { \ell + p - 1 \choose p - 1 }
-  \prod_{q=1}^{p-1} \p{\ell-q+F_{np;\mu}}.
-\end{align}
-Substitution of this result together with $D_{m-p,n-p;\mu}^{(\ell)}$
-using \eqref{eq:D_mn_k} into \eqref{eq:E_mn_p_sum} then gives us
-\begin{align}
-  E_{mn;\mu}^{(p)}
-  = \f{n!}{\p{n-p-1}!} S_\bmu^{n-p-1} S_\mu^{m-p-1} G_{mnp;\mu}
-\end{align}
-with
-\begin{align}
-  G_{mnp;\mu}
-  &\equiv \sum_{\ell=0}^{m-p-1} { \ell + p - 1 \choose p - 1 }
-  \p{m-p-\ell} \p{\ell+m-p+F_{np;\mu}}
-  \p{2\ell + F_{np;\mu}}
-  \prod_{q=1}^{p-1} \p{\ell-q+F_{np;\mu}} \\
-  &= { m \choose p + 1 } \prod_{q=0}^p \p{m-p-q+F_{np;\mu}}.
-\end{align}
-We can further simplify
-\begin{align}
-  \prod_{q=0}^p \p{m-p-q+F_{np;\mu}}
-  = \prod_{q=0}^p \p{m-n-q+2\mu S_\z}
-  = \sum_{q=0}^{p+1} \p{-1}^{p+1-q}
-  { p+1 \brack q } \p{m-n+2\mu S_\z}^q,
-\end{align}
-where ${ n \brack k }$ is an unsigned Stirling number of the first
-kind, and finally
-\begin{align}
-  \sum_{q=0}^p \p{-1}^{p-q} { p \brack q } \p{m-n+2\mu S_\z}^q
-  &= \sum_{q=0}^p \p{-1}^{p-q} { p \brack q } \sum_{\ell=0}^q
-  { q \choose \ell } \p{m-n}^{q-\ell} \p{2\mu S_\z}^\ell \\
-  &= \sum_{\ell=0}^p 2^\ell \sum_{q=\ell}^p \p{-1}^{p-q}
-  { p \brack q } { q \choose \ell } \p{m-n}^{q-\ell} \p{\mu S_\z}^\ell.
-\end{align}
-Putting everything together, we finally have
-\begin{align}
-  E_{mn;\mu}^{(p-1)}
-  = p! { m \choose p } { n \choose p }
-  S_\bmu^{n-p} S_\mu^{m-p}
-  \sum_{\ell=0}^p \epsilon_{mn}^{p\ell} \p{\mu S_\z}^\ell,
-\end{align}
-with
-\begin{align}
-  \epsilon_{mn}^{p\ell}
-  \equiv 2^\ell \sum_{q=\ell}^p \p{-1}^{p-q}
-  { p \brack q } { q \choose \ell } \p{m-n}^{q-\ell},
-\end{align}
-where in this final form $E_{mn;\mu}^{(0)} = D_{mn;\mu}$, which
-together with the expansion for $C_{mn;\mu}$ in \eqref{eq:C_mn_E}
-implies that
-\begin{align}
-  \sp{S_\mu^m, S_\bmu^n}_-
-  = \sum_{p=1}^{\min\set{m,n}}
-  p! { m \choose p } { n \choose p } S_\bmu^{n-p} S_\mu^{m-p}
-  \sum_{\ell=0}^p \epsilon_{mn}^{p\ell} \p{\mu S_\z}^\ell,
-  \label{eq:comm_mu}
-\end{align}
-and
-\begin{align}
-  S_\mu^m S_\bmu^n
-  = \sum_{p=0}^{\min\set{m,n}}
-  p! { m \choose p } { n \choose p } S_\bmu^{n-p} S_\mu^{m-p}
-  \sum_{\ell=0}^p \epsilon_{mn}^{p\ell} \p{\mu S_\z}^\ell,
-  \label{eq:push_mu_bmu}
-\end{align}
-If we wish to order products of collective spin operators with $S_\z$
-in between $S_\bmu$ and $S_\mu$, then
-\begin{align}
-  S_\mu^m S_\bmu^n
-  = \sum_{p=0}^{\min\set{m,n}} p! { m \choose p } { n \choose p }
-  S_\bmu^{n-p} Z_{mn;\bmu}^{(p)} S_\mu^{m-p},
-\end{align}
-where
-\begin{align}
-  Z_{mn;\bmu}^{(p)}
-  \equiv \sum_{\ell=0}^p \epsilon_{mn}^{p\ell}
-  \p{-\sp{m-p} + \mu S_\z}^\ell
-  = \sum_{\ell=0}^p \epsilon_{mn}^{p\ell}
-  \p{-1}^\ell \p{\sp{m-p} + \bmu S_\z}^\ell
-  = \sum_{q=0}^p \zeta_{mn}^{pq} \p{\bmu S_\z}^q,
-  \label{eq:Z_mnp}
-\end{align}
-with
-\begin{align}
-  \zeta_{mn}^{pq}
-  \equiv \sum_{\ell=q}^p \epsilon_{mn}^{p\ell} \p{-1}^\ell
-  { \ell \choose q } \p{m-p}^{\ell-q}
-  = \p{-1}^p 2^q \sum_{s=q}^p
-  { p \brack s } { s \choose q } \p{m+n-2p}^{s-q}.
-  \label{eq:zeta_mnpq}
-\end{align}
-
-\section{Product of arbitrary ordered collective spin operators}
-\label{sec:prod_general}
-
-The most general product of collective spin operators that we need to
-compute is
-\begin{align}
-  \S^{pqr}_{\ell mn;\mu}
-  = S_\mu^p \p{\mu S_\z}^q S_\bmu^r
-  S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n
-  = \sum_{k=0}^{\min\set{r,\ell}} k! { r \choose k } { \ell \choose k }
-  S_\mu^{p+\ell-k} \tilde Z_{qr\ell m;\mu}^{(k)} S_\bmu^{r+n-k},
-  \label{eq:general_product}
-\end{align}
-where
-\begin{align}
-  \tilde Z_{qr\ell m;\mu}^{(k)}
-  &\equiv \p{\ell-k+\mu S_\z}^q
-  Z_{r\ell;\mu}^{(k)} \p{r-k+\mu S_\z}^m \\
-  &= \sum_{a=0}^k \zeta_{r\ell}^{ka}
-  \sum_{b=0}^q \p{\ell-k}^{q-b} { q \choose b }
-  \sum_{c=0}^m \p{r-k}^{m-c} { m \choose c }
-  \p{\mu S_\z}^{a+b+c},
-\end{align}
-is defined in terms of $Z_{r\ell;\mu}^{(k)}$ and $\zeta_{r\ell}^{ka}$
-as respectively given in \eqref{eq:Z_mnp} and \eqref{eq:zeta_mnpq}.
-The (anti-)commutator of two ordered products of collective spin
-operators is then
-\begin{align}
-  \sp{S_\mu^p \p{\mu S_\z}^q S_\bmu^r,
-    S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_\pm
-  = \S^{pqr}_{\ell mn;\mu} \pm \S^{\ell mn}_{pqr;\mu}.
-\end{align}
-
-\section{Sandwich identities for single-spin decoherence calculations}
-\label{sec:sandwich_single}
-
-In this section we derive several identities which will be necessary
-for computing the effects of single-spin decoherence on ordered
-products of collective spin operators, i.e.~on operators of the form
-$S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$.  These identities all involve
-sandwiching a collective spin operator between operators that act on
-individual spins only, and summing over all individual spin indices.
-Our general strategy will be to use commutation relations to push
-single-spin operators together, and then evaluate the sum to arrive at
-an expression in terms of only collective spin operators.
-
-We first compute sums of single-spin operators sandwiching
-$\p{\mu S_\z}^m$, when necessary making use of the identity in
-\eqref{eq:push_z_mu_Ss}.  Up to Hermitian conjugation, the unique
-cases are, for $\mu,\nu\in\set{+1,-1}$,
-\begin{align}
-  \sum_j s_\z^{(j)} \p{\mu S_\z}^m s_\z^{(j)}
-  &= \sum_j s_\z^{(j)} s_\z^{(j)} \p{\mu S_\z}^m
-  = \f14 \sum_j \1_j \p{\mu S_\z}^m
-  = \f12 S \p{\mu S_\z}^m, \\
-  \sum_j s_\z^{(j)} \p{\mu S_\z}^m s_\nu^{(j)}
-  &= \p{\mu S_\z}^m \sum_j s_\z^{(j)} s_\nu^{(j)}
-  = \f12 \p{\mu S_\z}^m \nu S_\nu
-  = \f12 \nu S_\nu \p{\mu\nu+\mu S_\z}^m, \\
-  \sum_j s_\nu^{(j)} \p{\mu S_\z}^m s_\nu^{(j)}
-  &= \sum_j s_\nu^{(j)} s_\nu^{(j)} \p{\mu\nu+\mu S_\z}^m
-  = 0, \\
-  \sum_j s_\bnu^{(j)} \p{\mu S_\z}^m s_\nu^{(j)}
-  &= \sum_j s_\bnu^{(j)} s_\nu^{(j)} \p{\mu\nu+\mu S_\z}^m
-  = \p{S-\nu S_\z} \p{\mu\nu+\mu S_\z}^m.
-\end{align}
-We are now equipped to derive similar identities for more general
-collective spin operators.  Making heavy use of identities
-\eqref{eq:push_z_mu_sS} and \eqref{eq:push_mu_Ss} to push single-spin
-operators through transverse collective-spin operators, we again work
-through all combinations that are unique up to Hermitian conjugation,
-finding
-\begin{align}
-  \sum_j s_\z^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\z^{(j)}
-  &= \f12 \p{S-\ell-n} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n
-  + \ell n S_\mu^{\ell-1} \p{S+\mu S_\z}
-  \p{-1+\mu S_\z}^m S_\bmu^{n-1},
-  \label{eq:san_z_z} \allowdisplaybreaks \\
-  \sum_j s_\z^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\mu^{(j)}
-  &= \f12 \mu S_\mu^{\ell+1} \p{1+\mu S_\z}^m S_\bmu^n
-  - \mu n \p{S-\ell-\f12\sp{n-1}} S_\mu^\ell
-  \p{\mu S_\z}^m S_\bmu^{n-1} \notag \\
-  &\qquad - \mu\ell n\p{n-1} S_\mu^{\ell-1}
-  \p{S+\mu S_\z} \p{-1+\mu S_\z}^m S_\bmu^{n-2},
-  \label{eq:san_z_mu} \allowdisplaybreaks \\
-  \sum_j s_\z^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\bmu^{(j)}
-  &= -\f12 \mu S_\mu^\ell \p{\mu S_\z}^m S_\bmu^{n+1}
-  + \mu \ell S_\mu^{\ell-1} \p{S+\mu S_\z} \p{-1+\mu S_\z}^m S_\bmu^n,
-  \label{eq:san_z_bmu} \allowdisplaybreaks \\
-  \sum_j s_\mu^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\mu^{(j)}
-  &= n S_\mu^{\ell+1} \p{\mu S_\z}^m S_\bmu^{n-1}
-  - n\p{n-1} S_\mu^\ell \p{S+\mu S_\z} \p{-1+\mu S_\z}^m S_\bmu^{n-2},
-  \label{eq:san_mu_mu} \allowdisplaybreaks \\
-  \sum_j s_\mu^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\bmu^{(j)}
-  &= S_\mu^\ell \p{S+\mu S_\z}\p{-1+\mu S_\z}^m S_\bmu^n,
-  \label{eq:san_mu_bmu} \allowdisplaybreaks \\
-  \sum_j s_\bmu^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\mu^{(j)}
-  &= S_\mu^\ell \p{S - \ell - n - \mu S_\z}
-  \p{1+\mu S_\z}^m S_\bmu^n
-  + \ell n \p{2S - \ell - n + 2}
-  S_\mu^{\ell-1} \p{\mu S_\z}^m S_\bmu^{n-1} \notag \\
-  &\qquad + \ell n \p{\ell-1} \p{n-1} S_\mu^{\ell-2} \p{S+\mu S_\z}
-  \p{-1+\mu S_\z}^m S_\bmu^{n-2}.
-  \label{eq:san_bmu_mu}
-\end{align}
-
-\section{Uncorrelated, permutationally-symmetric single-spin
-  decoherence}
-\label{sec:decoherence_single}
-
-In this section we work out the effects of permutationally-symmetric
-decoherence of individual spins on collective spin operators of the
-form $S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$.  For shorthand, we define
-\begin{align}
-  \D\p{\gamma;\O} \equiv \D\p{\set{\gamma^{(j)}};\O}
-  = \sum_j\p{{\gamma^{(j)}}^\dag \O \gamma^{(j)}
-    - \f12\sp{{\gamma^{(j)}}^\dag \gamma^{(j)}, \O}_+},
-\end{align}
-where $\gamma$ is an operator that acts on a single spin, and
-$\gamma^{(j)}$ is an operator that acts with $\gamma$ on spin $j$ and
-the identity $\1$ on all other spins.
-
-\subsection{Decay-type decoherence}
-\label{sec:decay_single}
-
-The effect of decoherence via uncorrelated decay ($\mu=-1$) or
-excitation ($\mu=1$) of individual spins is described by
-\begin{align}
-  \D\p{s_\mu; \O}
-  = \sum_j\p{s_\bmu^{(j)} \O s_\mu^{(j)}
-    - \f12\sp{s_\bmu^{(j)} s_\mu^{(j)},\O}_+}
-  = \sum_j s_\bmu^{(j)} \O s_\mu^{(j)}
-  - S \O + \f{\mu}{2} \sp{S_\z, \O}_+.
-\end{align}
-In order to determine the effect of this decoherence on general
-collective spin operators, we expand the anti-commutator
-\begin{align}
-  \sp{S_\z, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
-  = S_\z S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n
-  + S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\z
-  = \mu S_\mu^\ell\p{\ell+n+2\mu S_\z} \p{\mu S_\z}^m S_\bmu^n,
-\end{align}
-which implies, using \eqref{eq:san_mu_bmu},
-\begin{align}
-  \D\p{s_\bmu; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
-  = S_\mu^\ell \p{S+\mu S_\z}\p{-1+\mu S_\z}^m S_\bmu^n
-  - S_\mu^\ell\sp{S + \f12\p{\ell+n} + \mu S_\z}
-  \p{\mu S_\z}^m S_\bmu^n,
-  \label{eq:decay_diff}
-\end{align}
-and, using \eqref{eq:san_bmu_mu},
-\begin{align}
-  \D\p{s_\mu; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
-  &= S_\mu^\ell \p{S - \ell - n - \mu S_\z} \p{1+\mu S_\z}^m S_\bmu^n
-  - S_\mu^\ell\sp{S - \f12\p{\ell+n} - \mu S_\z}
-  \p{\mu S_\z}^m S_\bmu^n \notag \\
-  &\qquad + \ell n \p{2S - \ell - n + 2}
-  S_\mu^{\ell-1} \p{\mu S_\z}^m S_\bmu^{n-1} \notag \\
-  &\qquad + \ell n \p{\ell-1} \p{n-1} S_\mu^{\ell-2} \p{S + \mu S_\z}
-  \p{-1+\mu S_\z}^m S_\bmu^{n-2}.
-  \label{eq:decay_same}
-\end{align}
-Decoherence via jump operators $s_\bmu^{(j)}$ only couples operators
-$S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$ to operators
-$S_\mu^\ell \p{\mu S_\z}^{m'} S_\bmu^n$ with $m'\le m$.  Decoherence
-via jump operators $s_\mu^{(j)}$, meanwhile, makes operators
-$S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$ ``grow'' in $m$ through the last
-term in \eqref{eq:decay_same}, although the sum $\ell+m+n$ does not
-grow.
-
-\subsection{Dephasing}
-\label{sec:dephasing_single}
-
-The effect of decoherence via single-spin dephasing is described by
-\begin{align}
-  \D\p{s_\z; \O}
-  = \sum_j\p{s_\z^{(j)} \O s_\z^{(j)}
-    - \f12\sp{s_\z^{(j)} s_\z^{(j)},\O}_+}
-  = \sum_j s_\z^{(j)} \O s_\z^{(j)} - \f12 S \O.
-\end{align}
-From \eqref{eq:san_z_z}, we then have
-\begin{align}
-  \D\p{s_\z; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
-  = -\f12\p{\ell+n} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n
-  + \ell n S_\mu^{\ell-1} \p{S + \mu S_\z}
-  \p{-1 + \mu S_\z}^m S_\bmu^{n-1}.
-\end{align}
-Decoherence via single-spin dephasing makes operators
-$S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$ ``grow'' in $m$, although the sum
-$\ell+m+n$ does not grow.
-
-\subsection{The general case}
-\label{sec:general_single}
-
-The most general type of single-spin decoherence is described by
-\begin{align}
-  \D\p{\gamma;\O}
-  = \sum_j\p{{\gamma^{(j)}}^\dag \O \gamma^{(j)}
-    - \f12\sp{{\gamma^{(j)}}^\dag \gamma^{(j)}, \O}_+},
-  &&
-  \gamma \equiv \gamma_\z s_\z + \gamma_+ s_+ + \gamma_- s_-.
-  \label{eq:D_general_single}
-\end{align}
-To simplify \eqref{eq:D_general_single}, we expand
-\begin{align}
-  \gamma^\dag \O \gamma
-  = \abs{\gamma_\z}^2 s_\z \O s_\z
-  + \sum_\mu \p{\abs{\gamma_\mu}^2 s_\bmu \O s_\mu
-    + \gamma_\bmu^* \gamma_\mu s_\mu \O s_\mu
-    + \gamma_\z^* \gamma_\mu s_\z \O s_\mu
-    + \gamma_\bmu^* \gamma_\z s_\mu \O s_\z},
-\end{align}
-and
-\begin{align}
-  \gamma^\dag \gamma
-  = \f14 \abs{\gamma_\z}^2
-  + \f12 \sum_\mu \sp{\abs{\gamma_\mu}^2 \p{1-2\mu s_\z}
-    + \mu \p{\gamma_\z^*\gamma_\mu - \gamma_\bmu^*\gamma_\z} s_\mu},
-\end{align}
-which implies
-\begin{align}
-  \D\p{\gamma;\O}
-  &= \sum_{X\in\set{\z,+,-}} \abs{\gamma_X}^2 \D\p{s_X;\O}
-  + \sum_{\mu,j}
-  \p{\gamma_\bmu^* \gamma_\mu s_\mu^{(j)} \O s_\mu^{(j)}
-    + \gamma_\z^* \gamma_\mu s_\z^{(j)} \O s_\mu^{(j)}
-    + \gamma_\bmu^* \gamma_\z s_\mu^{(j)} \O s_\z^{(j)}}
-  \notag \\
-  &\qquad -\f14 \sum_\mu \mu
-  \p{\gamma_\z^*\gamma_\mu - \gamma_\bmu^*\gamma_\z} \sp{S_\mu, \O}_+.
-\end{align}
-In order to compute the effect of this decoherence on general
-collective spin operators, we expand the anti-commutator
-\begin{align}
-  \sp{S_\mu, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
-  = S_\mu^{\ell+1} \sp{\p{\mu S_\z}^m+\p{1+\mu S_\z}^m} S_\bmu^n
-  - n S_\mu^\ell \p{n-1+2\mu S_\z} \p{\mu S_\z}^m S_\bmu^{n-1}.
-  \label{eq:S_mu_acomm}
-\end{align}
-Recognizing a resemblance between terms in \eqref{eq:S_mu_acomm} and
-\eqref{eq:san_z_mu}, we collect terms to simplify
-\begin{align}
-  \sum_j s_\z^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\mu^{(j)}
-  - \f14 \mu \sp{S_\mu, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
-  = K_{\ell mn;\mu} + L_{\ell mn;\mu}
-  \label{eq:dec_z_mu}
-\end{align}
-and likewise
-\begin{align}
-  \sum_j s_\mu^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\z^{(j)}
-  + \f14 \mu \sp{S_\mu, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
-  = K_{\ell mn;\mu} + M_{\ell mn;\mu}
-  \label{eq:dec_mu_z}
-\end{align}
-with
-\begin{align}
-  K_{\ell mn;\mu}
-  &\equiv \f14 \mu S_\mu^{\ell+1}
-  \sp{\p{1+\mu S_\z}^m-\p{\mu S_\z}^m} S_\bmu^n, \\
-  L_{\ell mn;\mu}
-  &\equiv -\mu n S_\mu^\ell \sp{S-\ell-\f34\p{n-1}-\f12\mu S_\z}
-  \p{\mu S_\z}^m S_\bmu^{n-1}
-  - \mu\ell n\p{n-1} S_\mu^{\ell-1}
-  \p{S+\mu S_\z} \p{-1+\mu S_\z}^m S_\bmu^{n-2}, \\
-  M_{\ell mn;\mu}
-  &\equiv \mu n S_\mu^\ell \sp{\p{S+\mu S_\z}\p{-1+\mu S_\z}^m
-    - \f12\p{\f12\sp{n-1}+\mu S_\z}\p{\mu S_\z}^m} S_\bmu^{n-1}.
-\end{align}
-Defining for completion
-\begin{align}
-  P_{\ell mn;\mu}
-  &\equiv \sum_j s_\mu^{(j)} S_\mu^\ell
-  \p{\mu S_\z}^m S_\bmu^n s_\mu^{(j)}
-  = n S_\mu^{\ell+1} \p{\mu S_\z}^m S_\bmu^{n-1}
-  - n\p{n-1} S_\mu^\ell \p{S+\mu S_\z} \p{-1+\mu S_\z}^m S_\bmu^{n-2},
-\end{align}
-and
-\begin{align}
-  Q_{\ell mn;\mu}^{(\gamma)}
-  \equiv \gamma_\bmu^* \gamma_\mu P_{\ell mn;\mu}
-  + \p{\gamma_\z^* \gamma_\mu + \gamma_\bmu^* \gamma_\z}
-  K_{\ell mn;\mu}
-  + \gamma_\z^* \gamma_\mu L_{\ell mn;\mu}
-  + \gamma_\bmu^* \gamma_\z M_{\ell mn;\mu},
-  \label{eq:Q_single}
-\end{align}
-we finally have
-\begin{align}
-  \D\p{\gamma; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
-  = \sum_{X\in\set{\z,+,-}} \abs{\gamma_X}^2
-  \D\p{s_X; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
-  + Q_{\ell mn;\mu}^{(\gamma)} + \sp{Q_{nm\ell;\mu}^{(\gamma)}}^\dag.
-\end{align}
-Note that the sum $\ell+m+n$ for operators
-$S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$ does not grow under this type of
-decoherence.
-
-\section{Sandwich identities for collective-spin decoherence
-  calculations}
-\label{sec:sandwich_collective}
-
-In analogy with the work in Appendix \ref{sec:sandwich_single}, in
-this section we work out sandwich identities necessary for
-collective-spin decoherence calculations.  The simplest cases are
-\begin{align}
-  S_\mu S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\bmu
-  &= S_\mu^{\ell+1} \p{\mu S_\z}^m S_\bmu^{n+1},
-  \allowdisplaybreaks \\
-  S_\mu S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\z
-  &= \mu S_\mu^{\ell+1} \p{n+\mu S_\z} \p{\mu S_\z}^m S_\bmu^n,
-  \allowdisplaybreaks \\
-  S_\z S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\z
-  &= S_\mu^\ell \sp{\ell n + \p{\ell+n} \mu S_\z + \p{\mu S_\z}^2}
-  \p{\mu S_\z}^m S_\bmu^n.
-\end{align}
-With a bit more work, we can also find
-\begin{align}
-  S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\mu
-  &= S_\mu^{\ell+1} \p{1+\mu S_\z}^m S_\bmu^n
-  - n S_\mu^\ell \p{n-1+2\mu S_\z} \p{\mu S_\z}^m S_\bmu^{n-1},
-\end{align}
-which implies
-\begin{align}
-  S_\mu S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\mu
-  &= S_\mu^{\ell+2} \p{1+\mu S_\z}^m S_\bmu^n
-  - n S_\mu^{\ell+1} \p{n-1+2\mu S_\z} \p{\mu S_\z}^m S_\bmu^{n-1},
-  \allowdisplaybreaks \\
-  S_\z S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\mu
-  &= \mu S_\mu^{\ell+1} \p{\ell+1+\mu S_\z} \p{1+\mu S_\z}^m S_\bmu^n
-  \notag \\
-  &\qquad - \mu n S_\mu^\ell
-  \sp{\ell\p{n-1} + \p{2\ell+n-1}\mu S_\z + 2\p{\mu S_\z}^2}
-  \p{\mu S_\z}^m S_\bmu^{n-1}.
-\end{align}
-Finally, we compute
-\begin{align}
-  S_\bmu S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\mu
-  &= \sp{S_\mu^\ell S_\bmu - \ell S_\mu^{\ell-1} \p{\ell-1+2\mu S_\z}}
-  \p{\mu S_\z}^m
-  \sp{S_\mu S_\bmu^n - n \p{n-1+2\mu S_\z} S_\bmu^{n-1}} \notag \\
-  &= S_\mu^\ell S_\bmu \p{\mu S_\z}^m S_\mu S_\bmu^n \notag \\
-  &\qquad - S_\mu^\ell
-  \sp{\ell\p{\ell+1} + n\p{n+1}+2\p{\ell+n}\mu S_\z}
-  \p{1+\mu S_\z}^m S_\bmu^n \notag \\
-  &\qquad + \ell n S_\mu^{\ell-1}
-  \sp{\p{\ell-1}\p{n-1}+2\p{\ell+n-2}\mu S_\z + 4\p{\mu S_\z}^2}
-  \p{\mu S_\z}^m S_\bmu^{n-1},
-\end{align}
-where
-\begin{multline}
-  S_\bmu \p{\mu S_\z}^m S_\mu
-  = S_\bmu S_\mu \p{1+\mu S_\z}^m
-  = \p{S_\mu S_\bmu - 2\mu S_\z} \p{1+\mu S_\z}^m
-  = S_\mu \p{2+\mu S_\z}^m S_\bmu - 2\mu S_\z \p{1+\mu S_\z}^m,
-\end{multline}
-so
-\begin{align}
-  S_\bmu S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\mu
-  &= S_\mu^{\ell+1} \p{2+\mu S_\z}^m S_\bmu^{n+1} \notag \\
-  &\qquad - S_\mu^\ell
-  \sp{\ell\p{\ell+1} + n\p{n+1}+2\p{\ell+n+1}\mu S_\z}
-  \p{1+\mu S_\z}^m S_\bmu^n \notag \\
-  &\qquad + \ell n S_\mu^{\ell-1}
-  \sp{\p{\ell-1}\p{n-1}+2\p{\ell+n-2}\mu S_\z + 4\p{\mu S_\z}^2}
-  \p{\mu S_\z}^m S_\bmu^{n-1}.
-\end{align}
-
-\section{Collective spin decoherence}
-\label{sec:decoherence_collective}
-
-In this section we work out the effects of collective decoherence on
-general collective spin operators.  For shorthand, we define
-\begin{align}
-  \D\p{\Gamma;\O}
-  \equiv \D\p{\set{\Gamma};\O}
-  = \Gamma^\dag \O \Gamma - \f12\sp{\Gamma^\dag \Gamma, \O}_+,
-\end{align}
-where $\Gamma$ is a collective spin jump operator.
-
-\subsection{Decay-type decoherence and dephasing}
-\label{sec:decay_dephasing_collective}
-
-Making use of the results in Appendix \ref{sec:sandwich_collective},
-we find that the effects of collective decay-type decoherence on
-general collective spin operators are given by
-\begin{align}
-  \D\p{S_\bmu; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
-  = -S_\mu^{\ell+1} \sp{\p{1+\mu S_\z}^m - \p{\mu S_\z}^m} S_\bmu^{n+1}
-  + \f12 S_\mu^\ell \sp{\ell\p{\ell-1} + n\p{n-1}
-    + 2\p{\ell+n}\mu S_\z} \p{\mu S_\z}^m S_\bmu^n,
-\end{align}
-and
-\begin{align}
-  \D\p{S_\mu; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
-  &= S_\mu^{\ell+1} \sp{\p{2+\mu S_\z}^m-\p{1+\mu S_\z}^m}
-  S_\bmu^{n+1} \notag \\
-  &\qquad - S_\mu^\ell
-  \sp{\ell\p{\ell+1} + n\p{n+1}+2\p{\ell+n+1}\mu S_\z}
-  \p{1+\mu S_\z}^m S_\bmu^n \notag \\
-  &\qquad + \f12 S_\mu^\ell
-  \sp{\ell\p{\ell+1} + n\p{n+1}+2\p{\ell+n+2}\mu S_\z}
-  \p{\mu S_\z}^m S_\bmu^n \notag \\
-  &\qquad + \ell n S_\mu^{\ell-1}
-  \sp{\p{\ell-1}\p{n-1}+2\p{\ell+n-2}\mu S_\z + 4\p{\mu S_\z}^2}
-  \p{\mu S_\z}^m S_\bmu^{n-1}.
-\end{align}
-Similarly, the effect of collective dephasing is given by
-\begin{align}
-  \D\p{S_\z; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
-  = -\f12 \p{\ell-n}^2 S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n.
-\end{align}
-
-\subsection{The general case}
-\label{sec:general_collective}
-
-More generally, we consider jump operators of the form
-\begin{align}
-  \Gamma \equiv \Gamma_\z S_\z + \Gamma_+ S_+ + \Gamma_- S_-,
-\end{align}
-whose decoherence effects are determined by
-\begin{align}
-  \Gamma^\dag \O \Gamma
-  = \abs{\Gamma_\z}^2 S_\z \O S_\z
-  + \sum_\mu \p{\abs{\Gamma_\mu}^2 S_\bmu \O S_\mu
-    + \Gamma_\bmu^* \Gamma_\mu S_\mu \O S_\mu
-    + \Gamma_\z^* \Gamma_\mu S_\z \O S_\mu
-    + \Gamma_\bmu^* \Gamma_\z S_\mu \O S_\z},
-\end{align}
-and
-\begin{align}
-  \Gamma^\dag \Gamma
-  = \abs{\Gamma_\z}^2 S_\z^2
-  + \sum_\mu \p{\abs{\Gamma_\mu}^2 S_\bmu S_\mu
-    + \Gamma_\z^*\Gamma_\mu S_\z S_\mu
-    + \Gamma_\bmu^* \Gamma_\z S_\mu S_\z
-    + \Gamma_\bmu^* \Gamma_\mu S_\mu^2},
-\end{align}
-which implies
-\begin{align}
-  \D\p{\Gamma;\O}
-  &= \sum_{X\in\set{\z,+,-}} \abs{\Gamma_X}^2 \D\p{S_X;\O}
-  + \sum_\mu \p{\Gamma_\bmu^* \Gamma_\mu S_\mu \O S_\mu
-    + \Gamma_\z^* \Gamma_\mu S_\z \O S_\mu
-    + \Gamma_\bmu^* \Gamma_\z S_\mu \O S_\z}
-  \notag \\
-  &\qquad -\f12 \sum_\mu\p{\Gamma_\bmu^* \Gamma_\mu \sp{S_\mu^2, \O}_+
-    + \Gamma_\z^*\Gamma_\mu \sp{S_\z S_\mu, \O}_+
-    + \Gamma_\bmu^* \Gamma_\z \sp{S_\mu S_\z, \O}_+}.
-\end{align}
-In order to compute the effect of this decoherence on general
-collective spin operators, we expand the anti-commutators
-\begin{align}
-  \sp{S_\mu^2, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
-  &= S_\mu^{\ell+2} \sp{\p{2+\mu S_\z}^m+\p{\mu S_\z}^m} S_\bmu^n
-  - 2n S_\mu^{\ell+1} \p{n+2\mu S_\z} \p{1+\mu S_\z}^m S_\bmu^{n-1}
-  \notag \\
-  &\qquad + n\p{n-1} S_\mu^\ell \sp{\p{n-1}\p{n-2}
-    + 2\p{2n-3}\mu S_\z + 4\p{\mu S_\z}^2} \p{\mu S_\z}^m S_\bmu^{n-2},
-  \allowdisplaybreaks \\
-  \sp{S_\z S_\mu, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
-  &= \mu S_\mu^{\ell+1} \sp{\p{\ell+1+\mu S_\z}\p{\mu S_\z}^m
-    + \p{n+1+\mu S_\z} \p{1+\mu S_\z}^m } S_\bmu^n \notag \\
-  &\qquad - \mu n S_\mu^\ell \sp{n \p{n-1}
-    + \p{3n-1}\mu S_\z + 2\p{\mu S_\z}^2} \p{\mu S_\z}^m S_\bmu^{n-1},
-  \allowdisplaybreaks \\
-  \sp{S_\mu S_\z, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
-  &= \mu S_\mu^{\ell+1} \sp{\p{\ell+\mu S_\z}\p{\mu S_\z}^m
-    + \p{n+\mu S_\z} \p{1+\mu S_\z}^m} S_\bmu^n \notag \\
-  &\qquad - \mu n S_\mu^\ell \sp{\p{n-1}^2
-    + 3\p{n-1}\mu S_\z + 2\p{\mu S_\z}^2} \p{\mu S_\z}^m S_\bmu^{n-1}.
-\end{align}
-Collecting terms and defining
-\begin{align}
-  \Gamma_{\z,\mu}^{(\pm)}
-  &\equiv \f12\p{\Gamma_\z^* \Gamma_\mu \pm \Gamma_\bmu^* \Gamma_\z},
-  \allowdisplaybreaks \\
-  \tilde L_{\ell mn;\mu}^{(\Gamma)}
-  &\equiv \mu \sp{\p{\ell-n+\f12} \Gamma_{\z,\mu}^{(+)}
-    + \p{\ell+\f12} \Gamma_{\z,\mu}^{(-)}}
-  S_\mu^{\ell+1} \p{1+\mu S_\z}^m S_\bmu^n \notag \\
-  &\qquad -\mu \sp{\p{\ell-n+\f12} \Gamma_{\z,\mu}^{(+)}
-    + \p{n+\f12} \Gamma_{\z,\mu}^{(-)}}
-  S_\mu^{\ell+1} \p{\mu S_\z}^m S_\bmu^n \notag \\
-  &\qquad + \mu \Gamma_{\z,\mu}^{(-)}
-  S_\mu^{\ell+1} \mu S_\z
-  \sp{\p{1+\mu S_\z}^m - \p{\mu S_\z}^m} S_\bmu^n,
-  \allowdisplaybreaks \\
-  \tilde M_{\ell mn;\mu}^{(\Gamma)}
-  &= -\mu n\p{n-1} \sp{\p{\ell-n+\f12} \Gamma_{\z,\mu}^{(+)}
-    + \p{\ell-\f12} \Gamma_{\z,\mu}^{(-)}}
-  S_\mu^\ell \p{\mu S_\z}^m S_\bmu^{n-1} \notag \\
-  &\qquad - 2\mu n \sp{\p{\ell-n+\f12} \Gamma_{\z,\mu}^{(+)}
-    + \p{\ell+\f12n-1} \Gamma_{\z,\mu}^{(-)}}
-  S_\mu^\ell \p{\mu S_\z}^{m+1} S_\bmu^{n-1} \notag \\
-  &\qquad - 2\mu n \Gamma_{\z,\mu}^{(-)}
-  S_\mu^\ell \p{\mu S_\z}^{m+2} S_\bmu^{n-1},
-  \allowdisplaybreaks \\
-  \tilde P_{\ell mn;\mu}
-  &\equiv -\f12 S_\mu^{\ell+2}
-  \sp{\p{2+\mu S_\z}^m - 2\p{1+\mu S_\z}^m + \p{\mu S_\z}^m}
-  S_\bmu^n \notag \\
-  &\qquad + n S_\mu^{\ell+1} \sp{\p{n+2\mu S_\z} \p{1+\mu S_\z}^m
-    - \p{n-1+2\mu S_\z} \p{\mu S_\z}^m}
-  S_\bmu^{n-1} \notag \\
-  &\qquad -n\p{n-1} S_\mu^\ell
-  \sp{\f12\p{n-1}\p{n-2} + \p{2n-3}\mu S_\z + 2\p{\mu S_\z}^2}
-  \p{\mu S_\z}^m S_\bmu^{n-2},
-  \allowdisplaybreaks \\
-  \tilde Q_{\ell mn;\mu}^{(\Gamma)}
-  &\equiv \Gamma_\bmu^* \Gamma_\mu \tilde P_{\ell mn;\mu}
-  + \tilde L_{\ell mn;\mu}^{(\Gamma)}
-  + \tilde M_{\ell mn;\mu}^{(\Gamma)},
-\end{align}
-we then have
-\begin{align}
-  \D\p{\Gamma; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
-  = \sum_{X\in\set{\z,+,-}} \abs{\Gamma_X}^2
-  \D\p{S_X; S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
-  + \tilde Q_{\ell mn;\mu}^{(\Gamma)}
-  + \sp{\tilde Q_{nm\ell;\mu}^{(\Gamma)}}^\dag.
-\end{align}
-Note that the sum $\ell+m+n$ for operators
-$S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$ grows by one if $\Gamma_\mu\ne0$
-or $\Gamma_\bmu\ne0$, and does not grow otherwise.
-
-\section{Initial conditions}
-\label{sec:initial_conditions}
-
-Initial conditions for all collective spin operators with respect to
-(spin-polarized) Gaussian states in the Dicke manifold spanned by
-$\set{\ket{m}:S_\z\ket{m}=m\ket{m}}$ can be determined from the
-expansions
-\begin{align}
-  S_\z = \sum_{k=-S}^S k \op{k},
-  &&
-  S_\mu = \sum_{k=-S+\delta_{\mu,-1}}^{S-\delta_{\mu,1}}
-  g_\mu\p{S,k} \op{k+\mu}{k},
-\end{align}
-with $\mu\in\set{+1,-1}$ and
-\begin{align}
-  g_\mu\p{S,k} \equiv \sqrt{\p{S-\mu k}\p{S+\mu k+1}}.
-\end{align}
-Given the initial states
-\begin{align}
-  \ket{\pm\Z} \equiv \ket{\pm S},
-  &&
-  \ket{\X} \equiv 2^{-N/2} \sum_{k=-S}^S
-  { N \choose S+k }^{1/2} \ket{k},
-\end{align}
-where $S_\x\ket{\X}=S\ket{\X}$, in terms of $\nu\in\set{+1,-1}$ and
-$\bmu\equiv-\mu$ we then have
-\begin{align}
-  \bk{\nu\Z|S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n|\nu\Z}
-  = \delta_{\ell n} \times
-  \begin{cases}
-     \p{S-n}^m \f{N! n!}{\p{N-n}!} & \mu = \nu, \\
-     \delta_{n,0} \p{-S}^m & \mu \ne \nu,
-  \end{cases}
-\end{align}
-and
-\begin{align}
-  \bk{\X|S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n|\X}
-  &=  \f1{2^N} \sum_{k=-S+\delta_{\mu,-1}\max\set{\ell,n}}
-  ^{S-\delta_{\mu,1}\max\set{\ell,n}} \p{\mu k}^m
-  \sp{{ N \choose S+k+\mu\ell } { N \choose S+k+\mu n }}^{1/2}
-  \notag \\
-  &\qquad\qquad\qquad\qquad
-  \times \prod_{p=0}^{\ell-1} g_\mu\p{S,k+\mu p}
-  \prod_{q=0}^{n-1} g_\bmu\p{S,k+\mu\sp{n-q}} \\
-  &= \f{N!}{2^N} \sum_{\mu k=-\mu S-\delta_{\mu,-1}\max\set{\ell,n}}
-  ^{\mu S-\delta_{\mu,1}\max\set{\ell,n}}
-  \f{\p{\mu k}^m \p{S-\mu k}!}
-  {\p{S+\mu k}!\p{S-\ell-\mu k}!\p{S-n-\mu k}!} \\
-  &= \f{N!}{2^N} \sum_{k=-S}^{S-\max\set{\ell,n}}
-  \f{k^m \p{S-k}!}{\p{S+k}!\p{S-\ell-k}!\p{S-n-k}!} \\
-  &= \p{-1}^m\f{N!}{2^N} \sum_{k=0}^{N-\max\set{\ell,n}}
-  \f{\p{S-k}^m \p{N-k}!}{k!\p{N-\ell-k}!\p{N-n-k}!}.
-\end{align}
-Furthermore, using the fact that
-\begin{align}
-  e^{i\phi S_\z} S_\mu e^{-i\phi S_\z} = e^{i\mu\phi} S_\mu,
-\end{align}
-we can define the state
-\begin{align}
-  \ket{\tilde\phi} \equiv e^{-i\phi S_\z}\ket{\X},
-  &&
-  \p{\cos\phi S_\x + \sin\phi S_\y} \ket{\tilde\phi}
-  = S \ket{\tilde\phi},
-\end{align}
-and compute
-\begin{align}
-  \bk{\tilde\phi|S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n|\tilde\phi}
-  = e^{i\mu\p{\ell-n}\phi} \bk{\X|S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n|\X}.
-\end{align}
-
-\section{Changing bases for collective spin operators}
+\section{Changing operator bases, and a general product for algebras
+  of permutationally-symmetric operators}
 \label{sec:new_basis}
@@
   \sum_{\substack{\p{\v j;\v m},\p{\v k;\v n} \\
-      \abs{\set{j_\alpha^\mu}\cap\set{k_\beta^\nu}}=s}} X.
+      \abs{\set{j_\a^\mu}\cap\set{k_\b^\nu}}=s}} X.
   \label{eq:SS_PP}
@@
   = m_\mu + n_\mu - \sum_\kappa \p{r_{\mu\kappa}+r_{\kappa\mu}}
-  + \sum_{\alpha,\beta} \rho^{\alpha\beta}_\mu.
+  + \sum_{\a,\b} \rho^{\a\b}_\mu.
 \end{align}
@@
 
-TODO: comment on the use of this new basis vs.~the old one.
-
-\bibliography{\jobname}
-
-\end{document}
+% TODO: comment on the use of this new basis vs.~the old one.

### 2019-09-11 soc_squeezing/collective_paper/collective_spin.tex
added comments on cumulant expansion added more discussion of the breakdown of TST added appendix on Heisenberg operators in open quantum systems, and the 'noise' operators

@@
-\documentclass[pra,twocolumn,longbibliography]{revtex4-2}
+\documentclass[aps,pra,twocolumn,longbibliography]{revtex4-2}
 
@@
 
+\renewcommand{\abs}[1]{\lvert #1 \rvert}
+
 \newcommand{\bk}{\braket} % shorthand for braket notation
@@
 
+\newcommand{\ul}{\underline} % shorthand for underline
+
 \newcommand{\A}{\mathcal{A}}
+\newcommand{\B}{\mathcal{B}}
 \newcommand{\D}{\mathcal{D}}
@@
 \renewcommand{\O}{\mathcal{O}}
+\newcommand{\Q}{\mathcal{Q}}
 \renewcommand{\S}{\mathcal{S}}
+\newcommand{\U}{\mathcal{U}}
 
@@
 for a range of applications including quantum-enhanced metrology and
-quantum simulation.  Theoretical interest in such systems dates back
-to the mid-twentieth century with the introduction of the Dicke
+quantum simulation.  The study of such systems dates back to the
+mid-twentieth century with the introduction of the Dicke
 model\cite{dicke1954coherence} that describes atoms cooperatively
@@
 collective spin models in a broad range of platforms, including atomic
-ensembles\cite{takano2009spin, appel2009mesoscopic, chen2010heralded,
+ensembles\cite{takano2009spin, appel2009mesoscopic,
   schleier-smith2010states, chen2011conditional}, Bose-Einstein
-condensates\cite{riedel2010atomchipbased, gross2010nonlinear,
-  klinder2015dynamical}, ultracold Fermi
-gasses\cite{martin2013quantum, smale2019observation}, trapped
-ions\cite{bohnet2016quantum}, optical
-cavities\cite{leroux2010implementation, bohnet2014reduced,
-  cox2016deterministic, hosten2016measurement, hosten2016quantum,
-  zhiqiang2017nonequilibrium, lewis-swan2018robust,
-  norcia2018cavitymediated}, and molecular
-sensors\cite{jones2009magnetic}, among others.  These implementations
-have triggered studies of a variety of rich subjects, including
-quantum criticality\cite{latorre2005entanglement,
-  alcalde2007functional, morrison2008dissipationdriven,
-  morrison2008collective, sarandy2009classical, wang2012quantum,
-  kessler2012dissipative, majd2014lmg}, non-equilibrium
+condensates\cite{esteve2008squeezing, riedel2010atomchipbased,
+  gross2010nonlinear, klinder2015dynamical}, ultracold Fermi
+gasses\cite{martin2013quantum}, trapped ions\cite{bohnet2016quantum},
+and optical cavities\cite{leroux2010implementation, bohnet2014reduced,
+  cox2016deterministic, hosten2016measurement,
+  norcia2018cavitymediated}, among others.  These implementations
+compliment theoretical studies in a variety of rich subjects,
+including quantum phase transitions and
+criticality\cite{latorre2005entanglement, alcalde2007functional,
+  wang2012quantum, majd2014lmg}, non-equilibrium
 phenomena\cite{walls1978nonequilibrium, morrison2008dynamical,
-  bhattacherjee2014nonequilibrium, klinder2015dynamical,
-  maghrebi2016nonequilibrium, zhiqiang2017nonequilibrium,
-  lang2018concurrence, smale2019observation}, and precision
-metrology\cite{wineland1992spin, itano1993quantum,
-  kitagawa1993squeezed, ma2011quantum, takano2009spin,
-  appel2009mesoscopic, chen2010heralded, schleier-smith2010states,
-  chen2011conditional, riedel2010atomchipbased, gross2010nonlinear,
-  zhong2010simplified, agarwal1997atomic, lau2014proposal,
-  huang2015quantum, wineland1992spin, kitagawa1993squeezed,
-  schleier-smith2010squeezing, leroux2010implementation,
-  bohnet2014reduced, cox2016deterministic, hosten2016measurement,
-  hosten2016quantum, lewis-swan2018robust, norcia2018cavitymediated,
-  jones2009magnetic, he2019engineering}.
+  morrison2008dissipationdriven, morrison2008collective,
+  kessler2012dissipative, bhattacherjee2014nonequilibrium,
+  zhiqiang2017nonequilibrium, lang2018concurrence}, and precision
+mentrology\cite{wineland1992spin, kitagawa1993squeezed,
+  zhong2010simplified, schleier-smith2010squeezing, ma2011quantum,
+  huang2015twoaxis, muessel2015twistandturn, huang2015quantum,
+  mirkhalaf2018robustifying, lewis-swan2018robust, he2019engineering}.
 
@@
 angle $\phi$ with $N$ independent spins to
-$\Delta\phi\sim1/\sqrt{N}$\cite{wineland1992spin, itano1993quantum,
-  ma2011quantum}.  Collective spin systems enable a means to break
-through this limit via the preparation of many-body entangled states
-such as NOON\cite{jones2009magnetic, chen2010heralded,
-  zhong2010simplified}, spin-cat\cite{agarwal1997atomic,
-  lau2014proposal, huang2015quantum}, and spin-squeezed
-states\cite{wineland1992spin, kitagawa1993squeezed, takano2009spin,
-  appel2009mesoscopic, schleier-smith2010states,
-  schleier-smith2010squeezing, leroux2010implementation,
-  riedel2010atomchipbased, gross2010nonlinear, chen2011conditional,
-  bohnet2014reduced, cox2016deterministic, hosten2016measurement,
-  hosten2016quantum, lewis-swan2018robust, norcia2018cavitymediated,
-  ma2011quantum, he2019engineering} that allow for measurement errors
-$\Delta\phi\sim1/N^\varepsilon$ with $1/2<\varepsilon\le1$, where
-$\varepsilon=1$ saturates the Heisenberg
-limit\cite{zwierz2010general}.  Generating such states typically
-requires nonlinear dynamics that are realized using collisional,
-photon-mediated, or phonon-mediated interactions.  Although a truly
-collective spin model requires uniform, all-to-all interactions, as
-long as measurements do not distinguish between constituent particles,
-even non-uniform systems may be effectively described by a uniform
-model with renormalized parameters\cite{hu2015entangled}.
+$\Delta\phi\sim1/\sqrt{N}$\cite{wineland1992spin, itano1993quantum}.
+Collective spin systems provide a means to break through this limit
+via the preparation of many-body entangled states such as spin-cat
+states\cite{agarwal1997atomic, lau2014proposal, huang2015quantum} and
+most notably spin-squeezed states\cite{wineland1992spin,
+  kitagawa1993squeezed, ma2011quantum} that allow for measurement
+errors $\Delta\phi\sim1/N^\varepsilon$ with $1/2<\varepsilon\le1$,
+where $\varepsilon=1$ saturates the Heisenberg
+limit\cite{zwierz2010general}.  Such entangled states can be prepared
+either via heralded methods such as quantum nondemolition
+measurements\cite{takano2009spin, appel2009mesoscopic,
+  schleier-smith2010states, chen2011conditional}, or via deterministic
+methods that require nonlinear dynamics, typically realized with
+phonon-mediated\cite{bohnet2016quantum},
+photon-mediated\cite{leroux2010implementation, bohnet2014reduced,
+  cox2016deterministic, hosten2016measurement,
+  norcia2018cavitymediated} or collisional\cite{esteve2008squeezing,
+  riedel2010atomchipbased, gross2010nonlinear, bromley2018dynamics}
+interactions.  Although a truly collective spin model requires
+uniform, all-to-all interactions, as long as measurements do not
+distinguish between constituent particles, even non-uniform systems
+may be effectively described by a uniform model with renormalized
+parameters\cite{hu2015entangled}.
 
@@
 requires the use of density operators, increasing the dimension of
-accessible state space to $\sim N^3$\cite{hartmann2016generalized,
+accessible state space to $O\p{N^3}$\cite{hartmann2016generalized,
   xu2013simulating}.  In this case, exact simulations can be carried
@@
 reproduce all expectation values of interest.  When decoherence is
-strong, however, or when the number of jump operators grows
-extensively with system size (e.g.~for single-spin decay), these Monte
-Carlo methods can take a prohibitively long time to converge, as
-simulations become dominated by incoherent jumps that generate large
-numbers of distinct quantum trajectories that need to be averaged in
-order to compute expectation values.
+strong, however, these Monte Carlo methods can take a prohibitively
+long time to converge, as simulations become dominated by incoherent
+jumps that generate large numbers of distinct quantum trajectories
+that need to be averaged in order to accurately compute expectation
+values.  Even with strong decoherence, dynamics are sometimes solvable
+through the cumulant expansion\cite{meiser2010steadystate} that
+neglects all $n$-body connected correlators for $n>2$.  The growth of
+genuinely multi-body correlations, however, eventually causes the
+cumulant expansion to yield unphysical results with no clear signature
+of failure.  In the absence of other means to compute correlators, it
+can therefore be difficult to identify the point at which correlators
+computed via the cumulant expansion can no longer be trusted.
 
@@
 of collective spin systems with generic Markovian decoherence.  The
-only restriction on decoherence is that, like the coherent collective
-dynamics, it must act identically on all constituent particles.  This
-method is based on a short-time expansion of exact solutions to the
-equations of motion for Heisenberg operators.  Evaluating this
-expansion requires knowing the structure constants of a collective
-spin operator algebra; the calculation of these structure constants
-(in Appendices \ref{sec:identities}--\ref{sec:general_product}) is one
-of the main technical results of this work, which we hope will empower
-both analytical and numerical studies of collective spin systems in
-the future.  We benchmark our method against both exact and quantum
+only restriction on decoherence (beyond Markovianity) is that, like
+the coherent collective dynamics, it must act identically on all
+constituent particles.  Our method is based on a short-time expansion
+of exact solutions to the equations of motion for Heisenberg
+operators.  Evaluating this expansion requires knowing the structure
+constants of a collective spin operator algebra; the calculation of
+these structure constants (in Appendices
+\ref{sec:identities}--\ref{sec:general_product}) is one of the main
+technical results of this work, which we hope will empower both
+analytical and numerical studies of collective spin systems in the
+future.  We benchmark our method against both exact and quantum
 trajectory Monte Carlo computations of spin squeezing in accessible
@@
 \section{Theory}
+\label{sec:theory}
 
@@
 expectation values of collective spin operators, deferring lengthy
-derivations to the appendices.  We consider systems composed of $N$
-distinct spin-1/2 particles.  Defining individual spin-1/2 operators
+derivations to the appendices.  We consider a system of $N$ distinct
+spin-1/2 particles.  Defining individual spin-1/2 operators
 $\hat s_{\a=\x,\y,\z}\equiv\hat\sigma_\a/2$ and
@@
 operators $\hat\E_\O\p{t}$ that result from interactions between the
-spin system and its environment, initially $\hat\E_\O\p{0}=0$.  The
-expectation values of Heisenberg operators evolve according to
+spin system and its environment, initially $\hat\E_\O\p{0}=0$.  These
+noise operators will essentially play no role in the present work, but
+are necessary to include for a consistent formalism of Heisenberg
+operators in an open quantum system; see Appendix \ref{sec:noise} for
+further discussion.  The expectation values of Heisenberg operators
+evolve according to
 %
@@
 \end{align}
-with a time derivative (super-)operator $\check T$ whose matrix
-elements $T_{\v m\v n}\in\C$ are defined by
+with a Heisenberg-picture time derivative operator $\check T=d/dt$
+whose matrix elements $T_{\v m\v n}\in\C$ are defined by
 \begin{align}
@@
 $\abs{\v\ell}\equiv\sum_\a\ell_\a$.  The growth of operator weight
-signifies the growth of many-body correlations.  In practice, keeping
-track of this growth will eventually require more computational
-resources than are available, meaning we must somehow truncate our
-equations of motion.  The simplest truncation strategy would be to
-take
+signifies the growth of many-body correlations.  Keeping track of this
+growth eventually becomes intractable, requiring us to truncate our
+equations of motion somehow.  The simplest truncation strategy would
+be to take
 \begin{align}
@@
 restrict the remainder of our discussions to the TST expansion in
-Eq.~\eqref{eq:TST}.
-%
-
-In principle, the factorial suppression of terms at higher orders of
-the expansion in Eq.~\eqref{eq:time_series} implies for any given time
-$t$, there exists a truncation order $M_t$ for which the truncation
-error in Eq.~\eqref{eq:TST} is negligibly small.  In practice, only a
-maximal truncation order $M_{\t{max}}$ is accessible with limited
-computational resources, such that the TST expansion in
-Eq.~\eqref{eq:TST} only allows us to compute the expectation value
-$\bk{\hat\O\p{t}}$ to a maximal time $t_{M_{\t{max}}}$.  As we will
-see in the following section, $M_{\t{max}}=35$ will suffice for the
-computation of collective spin correlators up to times that are
-relevant e.g.~for spin squeezing protocols.  Appendix
-\ref{sec:tutorial} provides a pedagogical tutorial for computing
-correlators using the TST expansion.
+Eq.~\eqref{eq:TST}, and provide a pedagogical tutorial for computing
+correlators using the TST expansion in Appendix \ref{sec:tutorial}.
+%
+
+%
+% In principle, the factorial suppression of terms at higher orders of
+% the expansion in Eq.~\eqref{eq:time_series} implies for any given
+% time $t$, there exists a truncation order $M_t$ for which the
+% truncation error in Eq.~\eqref{eq:TST} is negligibly small.  In
+% practice, only a maximal truncation order $M_{\t{max}}$ is
+% accessible with limited computational resources, such that the TST
+% expansion in Eq.~\eqref{eq:TST} only allows us to compute the
+% expectation value $\bk{\hat\O\p{t}}$ to a maximal time
+% $t_{M_{\t{max}}}$.  As we will see in the following section,
+% $M_{\t{max}}=35$ will suffice for the computation of collective spin
+% correlators up to times that are relevant e.g.~for spin squeezing
+% protocols.  Appendix \ref{sec:tutorial} provides a pedagogical
+% tutorial for computing correlators using the TST expansion.
+%
+Three primary considerations limit the maximum time $t$ to which we
+can accurately compute a correlator $\bk{\hat S_{\v n}\p{t}}$ using
+the TST expansion.  First, maintaining accuracy at larger times $t$
+requires going to higher orders $M$ in the TST expansion.  An
+order-$M$ TST expansion of the correlator $\bk{\hat S_{\v n}\p{t}}$
+can involve a significant fraction of operators $\hat S_{\v m}$ with
+weight $\abs{\v m}\lesssim M$, which implies the need to compute
+$O\p{M^3}$ initial-time expectation values $\bk{\hat S_{\v m}}$ and
+$O\p{M^4}$ matrix elements $T^k_{\v m\v n}$.  In practice, with a
+straightforward implementation of the TST expansion we find that these
+requirements generally restrict $M\lesssim 50$ -- $70$ with $8$ --
+$50$ gigabytes of random access memory (RAM).  Second, individual
+terms at high orders of the TST expansion in Eq.~\eqref{eq:TST} can
+grow excessively large, greatly amplifying any numerical errors and
+thereby spoiling cancellations that are necessary to arrive at a
+physical value of a correlator, i.e.~with
+$\abs{\bk{\hat S_{\v n}\p{t}}}\lesssim S^{\abs{\v n}}$.  Finally, the
+TST expansion is essentially perturbative in the time $t$, which
+implies that the validity of this a formal expansion eventually breaks
+down.  Precisely characterizing the implications of these last two
+considerations for the TST expansion requires additional analysis that
+we defer to future work.  As we show from benchmarks of the TST
+expansion in Section \ref{sec:squeezing}, however, a detailed
+understanding of these limitations is not necessary to diagnose the
+breakdown time $t_{\t{break}}^{(M)}$ beyond which the TST expansion
+yields inaccurate results.  Empirically, we find that going beyond
+order $M\approx35$ yields no significant gains in all cases considered
+in this work.
+%
 
@@
 twist-and-turn (TNT) models described by the collective spin
-Hamiltonians\cite{kitagawa1993squeezed, micheli2003manyparticle}
+Hamiltonians\cite{ma2011quantum}
 \begin{align}
@@
   \equiv \f{\p{\Delta\phi_{\t{min}}}^2}{\p{\Delta\phi_{\t{polarized}}}^2}
-  = \f{N}{\lvert\braket{\uv S}\rvert^2}
+  = \f{N}{\abs{\braket{\uv S}}^2}
   \min_{\substack{\v v\perp\bk{\uv S}\\\v v\cdot\v v=1}}
@@
 $\p{\hat s_\z,\hat s_\x}\to\p{\hat s_\x,-\hat s_\z}$ and
-$\ket\X\to\ket{-\Z}\equiv\ket\dn^{\otimes N}$ (together with
-appropriate transformations of the Hamiltonian and jump operators);
-the only effect of this rotation on the results presented in Figure
-\ref{fig:benchmarking} is to prolong the time for which the TST
-expansion agrees with the benchmarking simulations.  The reason for
-different results in rotated basis has to do with the breakdown of the
-TST expansion, which we discuss further below.
+$\ket\X\to\ket{-\Z}\equiv\ket\dn^{\otimes N}$, as well as appropriate
+transformations of the Hamiltonian and jump operators.
+%
+% The only effect of this rotation on the results presented in Figure
+% \ref{fig:benchmarking} is to prolong the time for which the TST
+% expansion agrees with the benchmarking simulations.  The reason for
+% different results in rotated basis has to do with the breakdown of
+% the TST expansion, which we discuss further below.
+%
+The only effects of this rotation on the results presented in Figure
+\ref{fig:benchmarking} are to
+\begin{enumerate*}
+\item reduce the time it takes to compute correlators
+  $\bk{\hat\O\p{t}}$ with the TST expansion, and
+\item prolong the time for which the TST expansion of TNT results
+  agree with benchmarking simulations.
+\end{enumerate*}
+The speedup in a different basis occurs because for the initial state
+$\ket{-\Z}$, all initial-time correlators $\bk{\hat S_{\v m}}$ are
+zero unless $m_+=m_-=0$, and all non-zero correlators take $O\p{1}$
+(i.e.~constant in $N$) time to compute, rather than $O\p{N}$ time (see
+Appendix \ref{sec:initial_conditions}).  In total, the use of a
+rotated basis reduces the computation time of initial-time correlators
+from $O\p{M^3N}$ to $O\p{M}$.  The reason for prolonged agreement of
+TNT results in a rotated basis is not entirely understood, and
+provides a clue into the precise mechanism by which the TST expansion
+breaks down (discussed below).  We defer a detailed study of this
+breakdown to future work.
+%
 
 The main lesson from Figure \ref{fig:benchmarking} is that the TST
-expansion gives essentially exact results right up until a sudden and
+expansion yields essentially exact results right up until a sudden and
 drastic departure that can be diagnosed by inspection.  The breakdown
@@
 unphysical squeezing parameter $\xi^2<0$, which occurs due to
-individual correlators taking unphysical values
-$\lvert\bk{\hat\S_{\v n}\p{t}}\rvert\gtrsim S^{\abs{\v n}}$.  The
-sudden and drastic departure from virtually exact results occurs
-because the TST expansion neglects high-weight
-(i.e.~large-$\abs{\v m}$) operators $\hat\S_{\v m}$ whose
-contributions to a Heisenberg operator $\hat\O\p{t}$ of interest
-eventually become non-negligible.  For sufficiently large times $t$,
-these individual contributions can become comparable in magnitude to
-the expectation value $\bk{\hat\O\p{t}}$, and will grow with some
-large power of $t$, so their truncation will have a drastic effect on
-the estimated value of $\bk{\hat\O\p{t}}$.  This breakdown mechanism
-is also the reason for prolonged agreement between the TST expansion
-and benchmarking methods when squeezing the initial state $\ket{-\Z}$
-rather than $\ket{\X}$: for initial states $\ket{-\Z}$, all
-initial-time correlators $\bk{\hat\S_{\v m}}$ vanish unless
-$m_+=m_-=0$ (see Appendix \ref{sec:initial_conditions}), so there is a
-substantially smaller number of neglected terms with non-zero
-contribution to the correlator $\bk{\hat\O\p{t}}$.
-
-\begin{figure}
-  \centering
-  \includegraphics{decoherence_strong.pdf}
-  \caption{Spin squeezing of $N=10^4$ spins initially in $\ket\X$ with
-    spontaneous decay, excitation, and dephasing of individual spins
-    at rates $\gamma_-=\gamma_+=\gamma_\z=100\chi$.  Computed using
-    the TST expansion in Eq.~\eqref{eq:TST} with $M=35$.  Solid
-    circles mark the times at which the TST expansion gives an
-    unphysical result with $\xi^2<0$.}
-  \label{fig:decoherence_strong}
-\end{figure}
+individual correlators taking unphysical values with
+$\abs{\bk{\hat\S_{\v n}\p{t}}}\gtrsim S^{\abs{\v n}}$.
+%
+% The sudden and drastic departure from virtually exact results occurs
+% because the TST expansion neglects high-weight
+% (i.e.~large-$\abs{\v m}$) operators $\hat\S_{\v m}$ whose
+% contributions to a Heisenberg operator $\hat\O\p{t}$ of interest
+% eventually become non-negligible.  For sufficiently large times $t$,
+% these individual contributions can become comparable in magnitude to
+% the expectation value $\bk{\hat\O\p{t}}$, and will grow with some
+% large power of $t$, so their truncation will have a drastic effect
+% on the estimated value of $\bk{\hat\O\p{t}}$.  This breakdown
+% mechanism is also the reason for prolonged agreement between the TST
+% expansion and benchmarking methods when squeezing the initial state
+% $\ket{-\Z}$ rather than $\ket{\X}$: for initial states $\ket{-\Z}$,
+% all initial-time correlators $\bk{\hat\S_{\v m}}$ vanish unless
+% $m_+=m_-=0$ (see Appendix \ref{sec:initial_conditions}), so there is
+% a substantially smaller number of neglected terms with non-zero
+% contribution to the correlator $\bk{\hat\O\p{t}}$.
+%
+Going up through order $M=70$ in the TST expansion does not
+significantly increase the breakdown time $t_{\t{break}}^{(M)}$, and
+in some cases even shortens $t_{\t{break}}^{(M)}$.  The sudden and
+drastic departure from virtually exact results is consistent with the
+limitations of the TST expansion discussed at the end of Section
+\ref{sec:theory}.  Specifically, we identify three possible mechanisms
+for breakdown:
+\begin{enumerate*}
+\item a rapid growth in the order $M$ necessary for the TST expansion
+  to converge,
+\item the growth of numerical errors in excessively large terms of the
+  TST expansion, and
+\item the formal breakdown of the perturbative expansion in the time
+  $t$.
+\end{enumerate*}
+In all of these cases, a detailed cancellation eventually ceases to
+occur between large terms at high orders in the TST expansion.  These
+large terms grow with the time $t$ raised to some large power (as high
+as $M$), and therefore rapidly yield wildly unphysical results.  In
+contrast to other approximate methods such as the cumulant
+expansion\cite{meiser2010steadystate}, the TST expansion can thus
+diagnose its own breakdown, which is an important feature when working
+in parameter regimes that are inaccessible via other means to compute
+correlators.
+%
 
@@
 
-Second, the TST expansion enables computing spin correlators in
-strong-decoherence regimes of large systems that are entirely
-inaccessible to other methods.  As an example, Figure
+\begin{figure}
+  \centering
+  \includegraphics{decoherence_strong.pdf}
+  \caption{Spin squeezing of $N=10^4$ spins initially in $\ket\X$ with
+    spontaneous decay, excitation, and dephasing of individual spins
+    at rates $\gamma_-=\gamma_+=\gamma_\z=100\chi$.  Computed using
+    the TST expansion in Eq.~\eqref{eq:TST} with $M=35$.  Solid
+    circles mark the times at which the TST expansion gives an
+    unphysical result with $\xi^2<0$.}
+  \label{fig:decoherence_strong}
+\end{figure}
+
+The second advantage of the TST expansion is the capability to compute
+spin correlators in strong-decoherence regimes of large systems that
+are entirely inaccessible to other methods.  As an example, Figure
 \ref{fig:decoherence_strong} shows squeezing of $N=10^4$ spins
@@
 $\hat\O\p{t}$.  While linear contributions from noise operators as
-e.g.~in Eq.~\eqref{eq:two_time} generally vanish, quadratic
-contributions that would otherwise appear in Eq.~\eqref{eq:four_point}
-generally do not.  We exclude these contributions because their
-calculation generally requires additional assumptions about the
-environment (i.e.~beyond Markovianity) that would be a digression for
-the purposes of the present work; see Ref.~\cite{blocher2019quantum}
-for a detailed discussion of noise terms and the quantum regression
-theorem underlying the calculation of multi-time correlators.
+e.g.~in Eq.~\eqref{eq:two_time} always vanish under Markovian
+decoherence (see Appendix \ref{sec:noise}), quadratic contributions
+that would otherwise appear in Eq.~\eqref{eq:four_point} generally do
+not\cite{blocher2019quantum}.  Determining the effect of these noise
+terms generally requires making additional assumptions about the
+environment, which would be a digression for the purposes of the
+present work.  We therefore exclude these noise terms in
+\eqref{eq:four_point} in order to keep our discussion simple and
+general; see Ref.~[\citenum{blocher2019quantum}] for more detailed
+discussions of noise terms and the quantum regression theorem
+underlying the calculation of multi-time correlators.
 %
@@
   garcia-mata2018chaos}.  In typical scenarios, such squared
-commutators initially vanish by construction due to a choice of
-operators with spatially separated non-trivial support.  Collective
-spin systems, however, have no intrinsic notion of locality or spatial
-separation.  In our case, therefore, with the choice of initial state
+commutators initially vanish by construction through to a choice of
+spatially separated operators.  Collective spin systems, however, have
+no intrinsic notion of locality or spatial separation.  In our case,
+therefore, with the choice of initial state
 $\ket\X\propto\p{\ket\up+\ket\dn}^{\otimes N}$ we merely have
@@
 
-\section{Analytical results for the one-axis twisting model}
-\label{sec:OAT}
-
-The one-axis twisting (OAT) Hamiltonian for $N$ spin-1/2 particles
-takes the form
-\begin{align}
-  H_{\t{OAT}}
-  = \chi S_\z^2
-  = \f12 \chi \sum_{j<k} \sigma_\z^{(j)} \sigma_\z^{(k)} + \f14 N \chi,
-\end{align}
-where $\sigma_\z^{(j)}$ represents a Pauli-$z$ operator acting on spin
-$j$.  This model is a special case of the zero-field Ising Hamiltonian
-previously solved in Ref.~[\citenum{foss-feig2013nonequilibrium}] via
-exact, analytical treatment of the quantum trajectory Monte Carlo
-method for computing expectation values.  The solution therein
-accounts for coherent evolution in addition to decoherence via
-uncorrelated single-spin decay, excitation, and dephasing respectively
-at rates $\gamma_-$, $\gamma_+$, and $\gamma_\z$ (denoted by
-$\Gamma_{\t{ud}}$, $\Gamma_{\t{du}}$, and $\Gamma_{\t{el}}$ in
-Ref.~[\citenum{foss-feig2013nonequilibrium}]).  Letting $S\equiv N/2$
-and $\mu,\nu\in\set{+1,-1}$, we adapt expectation values computed in
-Ref.~[\citenum{foss-feig2013nonequilibrium}] for the initial state
-$\ket\X\propto\p{\ket\up+\ket\dn}^{\otimes N}$ with
-$S_\x\ket\X=S\ket\X$ evolving under $H_{\t{OAT}}$, finding
-\begin{align}
-  \bk{S_+\p{t}}
-  &= S e^{-\Gamma t} \Phi\p{\chi,t}^{N-1}, \label{eq:S+_OAT} \\
-  \bk{\p{S_\mu S_\z}\p{t}}
-  &= -\f{\mu}{2}\bk{S_\mu\p{t}} + S \p{S-\f12} e^{-\Gamma t}
-  \Psi\p{\mu\chi,t} \Phi\p{\chi,t}^{N-2}, \\
-  \bk{\p{S_\mu S_\nu}\p{t}}
-  &= \delta_{\mu,-\nu} \p{S + \mu\bk{S_\z\p{t}}}
-  + S \p{S-\f12} e^{-2\Gamma t}
-  \Phi\p{\sp{\mu+\nu}\chi,t}^{N-2}, \label{eq:SS+-_OAT}
-\end{align}
-where
-\begin{align}
-  \Phi\p{X,t}
-  &\equiv e^{-\lambda t} \sp{\cos\p{t\sqrt{q_X^2-r}}
-    + \lambda t~\t{sinc}\p{t\sqrt{q_X^2-r}}},
-  \\
-  \Psi\p{X,t}
-  &\equiv e^{-\lambda t} \p{iq_X-\gamma}t~
-  \t{sinc}\p{t\sqrt{q_X^2-r}},
-\end{align}
-for
-\begin{align}
-  \gamma \equiv -\f12 \p{\gamma_+ - \gamma_-},
-  &&
-  \lambda \equiv \f12 \p{\gamma_+ + \gamma_-},
-  &&
-  \Gamma \equiv \lambda + \f12\gamma_\z,
-  &&
-  r \equiv \gamma_+ \gamma_-,
-  &&
-  q_X \equiv X + i\gamma,
-\end{align}
-In order to compute spin squeezing as measured by the Ramsey squeezing
-parameter $\xi^2$ defined in \eqref{eq:squeezing}, we additionally
-need analytical expressions for $\bk{S_\z\p{t}}$ and
-$\bk{S_\z^2\p{t}}$.  As these operators commute with both the OAT
-Hamiltonian and the single-spin operators $\sigma_\z^{(j)}$, their
-evolution is governed entirely by decay-type decoherence (see Appendix
-\ref{sec:decay_single}), which means
-\begin{align}
-  \f{d}{dt} \bk{S_\z\p{t}}
-  &= S\p{\gamma_+-\gamma_-} - \p{\gamma_++\gamma_-} \bk{S_\z\p{t}},
-  \\
-  \f{d}{dt} \bk{S_\z^2\p{t}}
-  &= S\p{\gamma_++\gamma_-}
-  + 2\p{S-\f12}\p{\gamma_+-\gamma_-} \bk{S_\z\p{t}}
-  - 2 \p{\gamma_++\gamma_-} \bk{S_\z^2\p{t}}.
-\end{align}
-The initial conditions $\bk{S_\z\p{0}}=0$ and $\bk{S_\z^2\p{0}}=S/2$
-then imply
-\begin{align}
-  \bk{S_\z\p{t}} = S\p{\f{\gamma_+-\gamma_-}{\gamma_++\gamma_-}}
-  \p{1-e^{-\p{\gamma_+ + \gamma_-} t}},
-  &&
-  \bk{S_\z^2\p{t}}
-  = S \sp{\f12 + \p{S-\f12} \p{\f{\bk{S_\z\p{t}}}{S}}^2}.
-  \label{eq:Sz_OAT}
-\end{align}
-With appropriate assumptions about the relevant sources of
-decoherence, the expectation values in
-\eqref{eq:S+_OAT}--\eqref{eq:SS+-_OAT} and \eqref{eq:Sz_OAT} are
-sufficient to compute the spin squeezing parameter $\xi^2$ in
-\eqref{eq:squeezing} at any time throughout evolution of the initial
-state $\ket\X$ under $H_{\t{OAT}}$.
-
-\section{Basic spin operator identities}
-\label{sec:identities}
-
-The appendices in this work make ubiquitous use of various spin
-operator identities; we collect and derive some basic identities here
-for reference.  Note that despite the working definition of collective
-spin operators from $S_\a=\sum_js_\a^{(j)}$, the identities we will
-derive involving only collective spin operators apply just as well to
-large-spin operators that cannot be expressed as the sum of individual
-spin-1/2 operators.  The elementary commutation relations between spin
-operators are, with $\bmu\equiv-\mu\in\set{+1,-1}$ for brevity,
-\begin{align}
-  \sp{s_\z^{(j)},s_\mu^{(k)}}_-
-  &= \delta_{jk} \mu s_\mu^{(j)},
-  &
-  \sp{S_\z,s_\mu^{(j)}}_-
-  &= \sp{s_\z^{(j)},S_\mu}_- = \mu s_\mu^{(j)},
-  &
-  \sp{S_\z,S_\mu}_-
-  &= \mu S_\mu,
-  \label{eq:comm_z_base} \\
-  \sp{s_\mu^{(j)},s_\bmu^{(k)}}_-
-  &= \delta_{jk} 2 \mu s_\z^{(j)},
-  &
-  \sp{S_\mu,s_\bmu^{(j)}}_-
-  &= \sp{s_\mu^{(j)},S_\bmu}_- = 2 \mu s_\z^{(j)},
-  &
-  \sp{S_\mu,S_\bmu}_-
-  &= 2 \mu S_\z.
-  \label{eq:comm_mu_base}
-\end{align}
-These relations can be used to inductively compute identities
-involving powers of collective spin operators.  By pushing through one
-spin operator at a time, we can find
-\begin{align}
-  \p{\mu S_\z}^m s_\mu^{(j)}
-  = \p{\mu S_\z}^{m-1} s_\mu^{(j)} \p{1 + \mu S_\z}
-  = \p{\mu S_\z}^{m-2} s_\mu^{(j)} \p{1 + \mu S_\z}^2
-  = \cdots
-  = s_\mu^{(j)} \p{1 + \mu S_\z}^m,
-  \label{eq:push_z_mu_Ss}
-\end{align}
-and
-\begin{align}
-  \mu s_\z^{(j)} S_\mu^m
-  = S_\mu \mu s_\z^{(j)} S_\mu^{m-1} + s_\mu^{(j)} S_\mu^{m-1}
-  = \cdots
-  = S_\mu^m \mu s_\z^{(j)} + ms_\mu^{(j)} S_\mu^{m-1},
-  \label{eq:push_z_mu_sS}
-\end{align}
-where we will generally find it nicer to express results in terms of
-$\mu s_\z^{(j)}$ and $\mu S_\z$ rather than $s_\z^{(j)}$ and $S_\z$.
-Summing over the single-spin index $j$ in both of the cases above
-gives us the purely collective-spin versions of these identities:
-\begin{align}
-  \p{\mu S_\z}^m S_\mu = S_\mu \p{1 + \mu S_\z}^m,
-  &&
-  \mu S_\z S_\mu^m = S_\mu^m \p{m + \mu S_\z},
-  \label{eq:push_z_mu_single}
-\end{align}
-where we can repeat the process of pushing through individual $S_\z$
-operators $\ell$ times to get
-\begin{align}
-  \p{\mu S_\z}^\ell S_\mu^m
-  = \p{\mu S_\z}^{\ell-1} S_\mu^m \p{m + \mu S_\z}
-  = \p{\mu S_\z}^{\ell-2} S_\mu^m \p{m + \mu S_\z}^2
-  = \cdots
-  = S_\mu^m \p{m + \mu S_\z}^\ell.
-  \label{eq:push_z_mu}
-\end{align}
-Multiplying \eqref{eq:push_z_mu} through by $\p{\mu\nu}^\ell$ (for
-$\nu\in\set{+1,-1}$) and taking its Hermitian conjugate, we can say
-that more generally
-\begin{align}
-  \p{\nu S_\z}^\ell S_\mu^m
-  = S_\mu^m \p{\mu\nu m+\nu S_\z}^\ell,
-  &&
-  S_\mu^m \p{\nu S_\z}^\ell
-  = \p{-\mu\nu m+\nu S_\z}^\ell S_\mu^m.
-\end{align}
-Finding commutation relations between powers of transverse spin
-operators, i.e.~$S_\mu$ and $S_\bmu$, turns out to be considerably
-more difficult than the cases we have worked out thus far.  We
-therefore save this work for Appendix \ref{sec:comm_transverse}.
-
-\section{Commutation relations between powers of transverse spin
-  operators}
-\label{sec:comm_transverse}
-
-To find commutation relations between powers of transverse collective
-spin operators, we first compute
-\begin{align}
-  S_\mu^m s_\bmu^{(j)}
-  &= S_\mu^{m-1}s_\bmu^{(j)} S_\mu
-  + S_\mu^{m-1} 2\mu s_\z^{(j)} \\
-  &= S_\mu^{m-2} s_\bmu^{(j)} S_\mu^2
-  + S_\mu^{m-2} 2\mu s_\z^{(j)} S_\mu
-  + S_\mu^{m-1} 2\mu s_\z^{(j)} \\
-  &= s_\bmu^{(j)} S_\mu^m
-  + \sum_{k=0}^{m-1} S_\mu^k 2\mu s_\z^{(j)} S_\mu^{m-k-1}
-  \label{eq:push_mu_Ss_start}.
-\end{align}
-While \eqref{eq:push_mu_Ss_start} gives us the commutator
-$\sp{S_\mu^m,s_\bmu^{(j)}}_-$, we would like to enforce an ordering on
-products of spin operators, which will ensure that we only keep track
-of operators that are linearly independent.  We choose (for now) to
-impose an ordering with all $s_\bmu^{(j)}$ operators on the left, and
-all $s_\z^{(j)}$ operators on the right.  Such an ordering will prove
-convenient for the calculations in this section\footnote{In
-  retrospect, it may have been nicer to push all $s_\mu^{(j)}$
-  operators to the right throughout these calculations, due to the
-  enhanced symmetry expressions would have with respect to Hermitian
-  conjugation.  In any case, we provide the final result of this
-  section in both ordering conventions, and therefore feel no need to
-  reproduce these calculations with a different ordering of spin
-  operators.}.  This choice of ordering compels us to expand
-\begin{align}
-  \sum_{k=0}^{m-1} S_\mu^k 2\mu s_\z^{(j)} S_\mu^{m-k-1}
-  &= \sum_{k=0}^{m-1} S_\mu^k
-  \sp{2\p{m-k-1} s_\mu^{(j)} S_\mu^{m-k-2}
-    + S_\mu^{m-k-1} 2\mu s_\z^{(j)}} \\
-  &= m \p{m-1} s_\mu^{(j)} S_\mu^{m-2}
-  + m S_\mu^{m-1} 2\mu s_\z^{(j)},
-\end{align}
-which implies
-\begin{align}
-  S_\mu^m s_\bmu^{(j)}
-  = s_\bmu^{(j)} S_\mu^m + m \p{m-1} s_\mu^{(j)} S_\mu^{m-2}
-  + m S_\mu^{m-1} 2\mu s_\z^{(j)},
-  \label{eq:push_mu_Ss}
-\end{align}
-and in turn
-\begin{align}
-  S_\mu^m S_\bmu = S_\bmu S_\mu^m
-  + m S_\mu^{m-1} \p{m - 1 + 2\mu S_\z}.
-  \label{eq:push_mu_single}
-\end{align}
-As the next logical step, we take on the task of computing
-\begin{align}
-  S_\mu^m S_\bmu^n
-  = S_\mu^{m-1} S_\bmu^n S_\mu
-  + n \sp{S_\mu^{m-1} S_\bmu^{n-1} \p{1 - n + 2\mu S_\z}}
-  = S_\bmu^n S_\mu^m
-  + n \sum_{k=0}^{m-1} S_\mu^{m-k-1} S_\bmu^{n-1}
-  \p{1 - n + 2\mu S_\z} S_\mu^k,
-\end{align}
-which implies
-\begin{align}
-  \sp{S_\mu^m, S_\bmu^n}_-
-  = C_{mn;\mu}
-  \equiv n \sum_{k=0}^{m-1} S_\mu^{m-k-1} S_\bmu^{n-1}
-  \p{1 - n + 2\mu S_\z} S_\mu^k.
-\end{align}
-We now need rearrange the operators in $C_{mn;\mu}$ into a standard
-order, which means pushing all $S_\z$ operators to the right and, for
-the purposes of this calculation, all $S_\bmu$ operators to the left.
-We begin by pushing $S_\mu^k$ to the left of $S_\z$, which takes
-$2\mu S_\z\to 2\mu S_\z+2k$, and then push $S_\mu^{m-k-1}$ to the
-right of $S_\bmu^{n-1}$, giving us
-\begin{align}
-  C_{mn;\mu}
-  &= n \sum_{k=0}^{m-1}
-  \p{S_\bmu^{n-1} S_\mu^{m-k-1} + C_{m-k-1,n-1;\mu}} S_\mu^k
-  \p{2k + 1 - n + 2\mu S_\z} \\
-  &= D_{mn;\mu}
-  + n \sum_{k=0}^{m-2} C_{m-k-1,n-1;\mu}
-  S_\mu^k \p{2k + 1 - n + 2\mu S_\z},
-  \label{eq:C_mn}
-\end{align}
-where we have dropped the last ($k=m-1$) term in the remaining sum
-because $C_{m-k-1,n-1;\mu}=0$ if $k=m-1$, and
-\begin{align}
-  D_{mn;\mu}
-  \equiv mn S_\bmu^{n-1} S_\mu^{m-1} \p{m - n + 2\mu S_\z}.
-  \label{eq:D_mn}
-\end{align}
-To our despair, we have arrived in \eqref{eq:C_mn} at a {\it
-  recursive} formula for $C_{mn;\mu}$.  Furthermore, we have not even
-managed to order all spin operators, as $C_{m-k-1,n-1;\mu}$ contains
-$S_\z$ operators that are to the left of $S_\mu^k$.  To sort all spin
-operators once and for all, we define
-\begin{align}
-  C_{mn;\mu}^{(k)} \equiv C_{m-k,n;\mu} S_\mu^k,
-  &&
-  D_{mn;\mu}^{(k)} \equiv D_{m-k,n;\mu} S_\mu^k,
-\end{align}
-which we can expand as
-\begin{align}
-  D_{mn;\mu}^{(k)}
-  &= \p{m-k}n S_\bmu^{n-1} S_\mu^{m-k-1}
-  \p{m-k-n+2\mu S_\z} S_\mu^k \\
-  &= \p{m-k}n S_\bmu^{n-1} S_\mu^{m-1} \p{k+m-n+2\mu S_\z},
-  \label{eq:D_mn_k}
-\end{align}
-and
-\begin{align}
-  C_{mn;\mu}^{(k)}
-  &= D_{m-k,n;\mu} S_\mu^k + n \sum_{j=0}^{m-k-2}
-  C_{m-k-j-1,n-1;\mu} S_\mu^j \p{2j+1-n+2\mu S_\z} S_\mu^k \\
-  &= D_{mn;\mu}^{(k)} + n \sum_{j=0}^{m-k-2}
-  C_{m-k-j-1,n-1;\mu} S_\mu^{j+k} \p{2j+2k+1-n+2\mu S_\z} \\
-  &= D_{mn;\mu}^{(k)} + n \sum_{j=0}^{m-k-2}
-  C_{m-1,n-1;\mu}^{(k+j)} \p{2\sp{j+k}+1-n+2\mu S_\z} \\
-  &= D_{mn;\mu}^{(k)} + n \sum_{j=k}^{m-2}
-  C_{m-1,n-1;\mu}^{(j)} \p{2j+1-n+2\mu S_\z}.
-  \label{eq:C_mn_k}
-\end{align}
-While the resulting expression in \eqref{eq:C_mn_k} strongly resembles
-that in \eqref{eq:C_mn}, there is one crucial difference: all spin
-operators in \eqref{eq:C_mn_k} have been sorted into a standard order.
-We can now repeatedly substitute $C_{mn;\mu}^{(k)}$ into itself, each
-time decreasing $m$ and $n$ by 1, until one of $m$ or $n$ reaches
-zero.  Such repeated substitution yields the expansion
-\begin{align}
-  C_{mn;\mu}
-  = C_{mn;\mu}^{(0)}
-  = D_{mn;\mu}
-  + \sum_{p=1}^{\min\set{m,n}-1} E_{mn;\mu}^{(p)},
-  \label{eq:C_mn_E}
-\end{align}
-where the first two terms in the sum over $p$ are
-\begin{align}
-  E_{mn;\mu}^{(1)}
-  &= n \sum_{k=0}^{m-2} D_{m-1,n-1;\mu}^{(k)} \p{2k+1-n+2\mu S_\z}, \\
-  E_{mn;\mu}^{(2)}
-  &= n \sum_{k_1=0}^{m-2} \p{n-1} \sum_{k_2=k_1}^{m-3}
-  D_{m-2,n-2;\mu}^{(k_2)} \p{2k_2+2-n+2\mu S_\z} \p{2k_1+1-n+2\mu S_\z},
-\end{align}
-and more generally for $p>1$,
-\begin{align}
-  E_{mn;\mu}^{(p)}
-  = \f{n!}{\p{n-p}!}
-  \sum_{k_1=0}^{m-2} \sum_{k_2=k_1}^{m-3} \cdots\sum_{k_p=k_{p-1}}^{m-p-1}
-  D_{m-p,n-p;\mu}^{(k_p)} \prod_{j=1}^p \p{2k_j+j-n+2\mu S_\z}.
-  \label{eq:E_mn_p}
-\end{align}
-In principle, the expressions in \eqref{eq:D_mn}, \eqref{eq:D_mn_k},
-\eqref{eq:C_mn_E}, and \eqref{eq:E_mn_p} suffice to evaluate the
-commutator $\sp{S_\mu^m,S_\bmu^n}_- = C_{mn;\mu}$, but this result is
--- put lightly -- quite a mess: the expression for $E_{mn;\mu}^{(p)}$
-in \eqref{eq:E_mn_p} involves a sum over $p$ mutually dependent
-intermediate variables, each term of which additionally contains a
-product of $p$ factors.  We therefore devote the rest of this section
-to simplifying our result for the commutator
-$\sp{S_\mu^m,S_\bmu^n}_-$.
-
-Observing that in \eqref{eq:E_mn_p} we always have
-$0\le k_1\le k_2\le\cdots\le k_p\le m-p-1$, we can rearrange the order
-of the sums and relabel $k_p\to\ell$ to get
-\begin{align}
-  E_{mn;\mu}^{(p)}
-  = \f{n!}{\p{n-p}!}
-  \sum_{\ell=0}^{m-p-1} D_{m-p,n-p;\mu}^{(\ell)} \p{2\ell+F_{np;\mu}}
-  \sum_{\p{\v k,p-1,\ell}} \prod_{j=1}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}},
-  \label{eq:E_mn_p_sum}
-\end{align}
-where for shorthand we define
-\begin{align}
-  F_{np;\mu} \equiv p - n + 2\mu S_\z,
-  &&
-  \sum_{\p{\v k,q,\ell}} X \equiv
-  \sum_{k_1=0}^\ell \sum_{k_2=k_1}^\ell
-  \cdots \sum_{k_q=k_{q-1}}^\ell X.
-\end{align}
-We now further define
-\begin{align}
-  f_{np\ell;\mu}\p{k,q} \equiv \p{\ell-k+q} \p{\ell+k-q+F_{np;\mu}},
-\end{align}
-and evaluate sums successively over $k_{p-1},k_{p-2},\cdots,k_1$,
-finding
-\begin{align}
-  \sum_{\p{\v k,p-1,\ell}} \prod_{j=1}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}}
-  &= \sum_{\p{\v k,p-2,\ell}}
-  \prod_{j=2}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}}
-  f_{np\ell;\mu}\p{k_{p-2},1} \\
-  &= \f1{\p{r-1}!} \sum_{\p{\v k,p-r,\ell}}
-  \prod_{j=r}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}}
-  \prod_{q=1}^{r-1} f_{np\ell;\mu}\p{k_{p-r},q} \\
-  &= \f1{\p{p-1}!} \prod_{q=1}^{p-1} f_{np\ell;\mu}\p{0,q} \\
-  &= { \ell + p - 1 \choose p - 1 }
-  \prod_{q=1}^{p-1} \p{\ell-q+F_{np;\mu}}.
-\end{align}
-Substitution of this result together with $D_{m-p,n-p;\mu}^{(\ell)}$
-using \eqref{eq:D_mn_k} into \eqref{eq:E_mn_p_sum} then gives us
-\begin{align}
-  E_{mn;\mu}^{(p)}
-  = \f{n!}{\p{n-p-1}!} S_\bmu^{n-p-1} S_\mu^{m-p-1} G_{mnp;\mu}
-\end{align}
-with
-\begin{align}
-  G_{mnp;\mu}
-  &\equiv \sum_{\ell=0}^{m-p-1} { \ell + p - 1 \choose p - 1 }
-  \p{m-p-\ell} \p{\ell+m-p+F_{np;\mu}}
-  \p{2\ell + F_{np;\mu}}
-  \prod_{q=1}^{p-1} \p{\ell-q+F_{np;\mu}} \\
-  &= { m \choose p + 1 } \prod_{q=0}^p \p{m-p-q+F_{np;\mu}}.
-\end{align}
-We can further simplify
-\begin{align}
-  \prod_{q=0}^p \p{m-p-q+F_{np;\mu}}
-  = \prod_{q=0}^p \p{m-n-q+2\mu S_\z}
-  = \sum_{q=0}^{p+1} \p{-1}^{p+1-q}
-  { p+1 \brack q } \p{m-n+2\mu S_\z}^q,
-\end{align}
-where ${ p \brack q }$ is an unsigned Stirling number of the first
-kind, and finally
-\begin{align}
-  \sum_{q=0}^p \p{-1}^{p-q} { p \brack q } \p{m-n+2\mu S_\z}^q
-  &= \sum_{q=0}^p \p{-1}^{p-q} { p \brack q } \sum_{\ell=0}^q
-  { q \choose \ell } \p{m-n}^{q-\ell} \p{2\mu S_\z}^\ell \\
-  &= \sum_{\ell=0}^p 2^\ell \sum_{q=\ell}^p \p{-1}^{p-q}
-  { p \brack q } { q \choose \ell } \p{m-n}^{q-\ell} \p{\mu S_\z}^\ell.
-\end{align}
-Putting everything together, we finally have
-\begin{align}
-  E_{mn;\mu}^{(p-1)}
-  = p! { m \choose p } { n \choose p }
-  S_\bmu^{n-p} S_\mu^{m-p}
-  \sum_{\ell=0}^p \epsilon_{mn}^{p\ell} \p{\mu S_\z}^\ell,
-\end{align}
-with
-\begin{align}
-  \epsilon_{mn}^{p\ell}
-  \equiv 2^\ell \sum_{q=\ell}^p \p{-1}^{p-q}
-  { p \brack q } { q \choose \ell } \p{m-n}^{q-\ell},
-\end{align}
-where in this final form $E_{mn;\mu}^{(0)} = D_{mn;\mu}$, which
-together with the expansion for $C_{mn;\mu}$ in \eqref{eq:C_mn_E}
-implies that
-\begin{align}
-  \sp{S_\mu^m, S_\bmu^n}_-
-  = \sum_{p=1}^{\min\set{m,n}}
-  p! { m \choose p } { n \choose p } S_\bmu^{n-p} S_\mu^{m-p}
-  \sum_{\ell=0}^p \epsilon_{mn}^{p\ell} \p{\mu S_\z}^\ell,
-  \label{eq:comm_mu}
-\end{align}
-and
-\begin{align}
-  S_\mu^m S_\bmu^n
-  = \sum_{p=0}^{\min\set{m,n}}
-  p! { m \choose p } { n \choose p } S_\bmu^{n-p} S_\mu^{m-p}
-  \sum_{\ell=0}^p \epsilon_{mn}^{p\ell} \p{\mu S_\z}^\ell.
-  \label{eq:push_mu_bmu}
-\end{align}
-If we wish to order products of collective spin operators with $S_\z$
-in between $S_\bmu$ and $S_\mu$, then
-\begin{align}
-  S_\mu^m S_\bmu^n
-  = \sum_{p=0}^{\min\set{m,n}} p! { m \choose p } { n \choose p }
-  S_\bmu^{n-p} Z_{mn;\bmu}^{(p)} S_\mu^{m-p},
-\end{align}
-where
-\begin{align}
-  Z_{mn;\bmu}^{(p)}
-  \equiv \sum_{\ell=0}^p \epsilon_{mn}^{p\ell}
-  \p{-\sp{m-p} + \mu S_\z}^\ell
-  = \sum_{q=0}^p \zeta_{mn}^{pq} \p{\bmu S_\z}^q,
-  \label{eq:Z_mnp}
-\end{align}
-with
-\begin{align}
-  \zeta_{mn}^{pq}
-  \equiv \sum_{\ell=q}^p \epsilon_{mn}^{p\ell}
-  { \ell \choose q } \p{-1}^\ell \p{m-p}^{\ell-q}
-  = \p{-1}^p 2^q \sum_{s=q}^p
-  { p \brack s } { s \choose q } \p{m+n-2p}^{s-q}.
-  \label{eq:zeta_mnpq}
-\end{align}
-Here ${ p \brack s }$ is an unsigned Stirling number of the first
-kind.
-
-\section{Product of arbitrary ordered collective spin operators}
-\label{sec:general_product}
-
-The most general product of collective spin operators that we need to
-compute is
-\begin{align}
-  \S^{pqr}_{\ell mn;\mu}
-  = S_\mu^p \p{\mu S_\z}^q S_\bmu^r
-  S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n
-  = \sum_{k=0}^{\min\set{r,\ell}} k! { r \choose k } { \ell \choose k }
-  S_\mu^{p+\ell-k} \tilde Z_{qr\ell m;\mu}^{(k)} S_\bmu^{r+n-k},
-  \label{eq:general_product}
-\end{align}
-where
-\begin{align}
-  \tilde Z_{qr\ell m;\mu}^{(k)}
-  &\equiv \p{\ell-k+\mu S_\z}^q
-  Z_{r\ell;\mu}^{(k)} \p{r-k+\mu S_\z}^m \\
-  &= \sum_{a=0}^k \zeta_{r\ell}^{ka}
-  \sum_{b=0}^q \p{\ell-k}^{q-b} { q \choose b }
-  \sum_{c=0}^m \p{r-k}^{m-c} { m \choose c }
-  \p{\mu S_\z}^{a+b+c},
-\end{align}
-is defined in terms of $Z_{r\ell;\mu}^{(k)}$ and $\zeta_{r\ell}^{ka}$
-as respectively given in \eqref{eq:Z_mnp} and \eqref{eq:zeta_mnpq}.
-The (anti-)commutator of two ordered products of collective spin
-operators is then simply
-\begin{align}
-  \sp{S_\mu^p \p{\mu S_\z}^q S_\bmu^r,
-    S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_\pm
-  = \S^{pqr}_{\ell mn;\mu} \pm \S^{\ell mn}_{pqr;\mu}.
-\end{align}
-
-\section{Sandwich identities for single-spin decoherence calculations}
-\label{sec:sandwich_single}
-
-In this section we derive several identities that will be necessary
-for computing the effects of single-spin decoherence on ordered
-products of collective spin operators, i.e.~on operators of the form
-$S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$.  These identities all involve
-sandwiching a collective spin operator between operators that act on
-individual spins only, and summing over all individual spin indices.
-Our general strategy will be to use commutation relations to push
-single-spin operators together, and then evaluate the sum to arrive at
-an expression involving only collective spin operators.
-
-We first compute sums of single-spin operators sandwiching
-$\p{\mu S_\z}^m$, when necessary making use of the identity in
-\eqref{eq:push_z_mu_Ss}.  Up to Hermitian conjugation, the unique
-cases are, for $S\equiv N/2$ and $\mu,\nu\in\set{+1,-1}$,
-\begin{align}
-  \sum_j s_\z^{(j)} \p{\mu S_\z}^m s_\z^{(j)}
-  &= \sum_j s_\z^{(j)} s_\z^{(j)} \p{\mu S_\z}^m
-  = \f14 \sum_j \1_j \p{\mu S_\z}^m
-  = \f12 S \p{\mu S_\z}^m, \\
-  \sum_j s_\z^{(j)} \p{\mu S_\z}^m s_\nu^{(j)}
-  &= \p{\mu S_\z}^m \sum_j s_\z^{(j)} s_\nu^{(j)}
-  = \f12 \p{\mu S_\z}^m \nu S_\nu
-  = \f12 \nu S_\nu \p{\mu\nu+\mu S_\z}^m, \\
-  \sum_j s_\nu^{(j)} \p{\mu S_\z}^m s_\nu^{(j)}
-  &= \sum_j s_\nu^{(j)} s_\nu^{(j)} \p{\mu\nu+\mu S_\z}^m
-  = 0, \\
-  \sum_j s_\bnu^{(j)} \p{\mu S_\z}^m s_\nu^{(j)}
-  &= \sum_j s_\bnu^{(j)} s_\nu^{(j)} \p{\mu\nu+\mu S_\z}^m
-  = \p{S-\nu S_\z} \p{\mu\nu+\mu S_\z}^m.
-\end{align}
-We are now equipped to derive similar identities for more general
-collective spin operators.  Making heavy use of identities
-\eqref{eq:push_z_mu_sS} and \eqref{eq:push_mu_Ss} to push single-spin
-operators through transverse collective-spin operators, we again work
-through all combinations that are unique up to Hermitian conjugation,
-finding
-\begin{align}
-  \sum_j s_\z^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\z^{(j)}
-  &= \f12 \p{S-\ell-n} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n
-  + \ell n S_\mu^{\ell-1} \p{S+\mu S_\z}
-  \p{-1+\mu S_\z}^m S_\bmu^{n-1},
-  \label{eq:san_z_z} \allowdisplaybreaks \\
-  \sum_j s_\z^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\mu^{(j)}
-  &= \f12 \mu S_\mu^{\ell+1} \p{1+\mu S_\z}^m S_\bmu^n
-  - \mu n \p{S-\ell-\f12\sp{n-1}} S_\mu^\ell
-  \p{\mu S_\z}^m S_\bmu^{n-1} \notag \\
-  &\qquad - \mu\ell n\p{n-1} S_\mu^{\ell-1}
-  \p{S+\mu S_\z} \p{-1+\mu S_\z}^m S_\bmu^{n-2},
-  \label{eq:san_z_mu} \allowdisplaybreaks \\
-  \sum_j s_\z^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\bmu^{(j)}
-  &= -\f12 \mu S_\mu^\ell \p{\mu S_\z}^m S_\bmu^{n+1}
-  + \mu \ell S_\mu^{\ell-1} \p{S+\mu S_\z} \p{-1+\mu S_\z}^m S_\bmu^n,
-  \label{eq:san_z_bmu} \allowdisplaybreaks \\
-  \sum_j s_\mu^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\mu^{(j)}
-  &= n S_\mu^{\ell+1} \p{\mu S_\z}^m S_\bmu^{n-1}
-  - n\p{n-1} S_\mu^\ell \p{S+\mu S_\z} \p{-1+\mu S_\z}^m S_\bmu^{n-2},
-  \label{eq:san_mu_mu} \allowdisplaybreaks \\
-  \sum_j s_\mu^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\bmu^{(j)}
-  &= S_\mu^\ell \p{S+\mu S_\z}\p{-1+\mu S_\z}^m S_\bmu^n,
-  \label{eq:san_mu_bmu} \allowdisplaybreaks \\
-  \sum_j s_\bmu^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\mu^{(j)}
-  &= S_\mu^\ell \p{S - \ell - n - \mu S_\z}
-  \p{1+\mu S_\z}^m S_\bmu^n
-  + \ell n \p{2S - \ell - n + 2}
-  S_\mu^{\ell-1} \p{\mu S_\z}^m S_\bmu^{n-1} \notag \\
-  &\qquad + \ell n \p{\ell-1} \p{n-1} S_\mu^{\ell-2} \p{S+\mu S_\z}
-  \p{-1+\mu S_\z}^m S_\bmu^{n-2}.
-  \label{eq:san_bmu_mu}
-\end{align}
-
-\section{Uncorrelated, permutationally-symmetric single-spin
-  decoherence}
-\label{sec:decoherence_single}
-
-In this section we work out the effects of permutationally-symmetric
-decoherence of individual spins on collective spin operators of the
-form $S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$.  For compactness, we define
-\begin{align}
-  \D\p{g} \O
-  \equiv \D\p{\set{g^{(j)}:j=1,2,\cdots,N}} \O
-  = \sum_j\p{{g^{(j)}}^\dag \O g^{(j)}
-    - \f12\sp{{g^{(j)}}^\dag g^{(j)}, \O}_+},
-\end{align}
-where $g$ is an operator that acts on a single spin, $g^{(j)}$ is an
-operator that acts with $g$ on spin $j$ and trivially on all other
-spins, and $N$ is the total number of spins.
-
-\subsection{Decay-type decoherence}
-\label{sec:decay_single}
-
-The effect of decoherence via uncorrelated decay ($\mu=-1$) or
-excitation ($\mu=1$) of individual spins is described by
-\begin{align}
-  \D\p{s_\mu} \O
-  = \sum_j\p{s_\bmu^{(j)} \O s_\mu^{(j)}
-    - \f12\sp{s_\bmu^{(j)} s_\mu^{(j)},\O}_+}
-  = \sum_j s_\bmu^{(j)} \O s_\mu^{(j)}
-  - S \O + \f{\mu}{2} \sp{S_\z, \O}_+.
-\end{align}
-In order to determine the effect of this decoherence on general
-collective spin operators, we expand the anti-commutator
-\begin{align}
-  \sp{S_\z, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
-  = S_\z S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n
-  + S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\z
-  = \mu S_\mu^\ell\p{\ell+n+2\mu S_\z} \p{\mu S_\z}^m S_\bmu^n,
-\end{align}
-which implies, using \eqref{eq:san_mu_bmu},
-\begin{align}
-  \D\p{s_\bmu} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
-  = S_\mu^\ell \p{S+\mu S_\z}\p{-1+\mu S_\z}^m S_\bmu^n
-  - S_\mu^\ell\sp{S + \f12\p{\ell+n} + \mu S_\z}
-  \p{\mu S_\z}^m S_\bmu^n,
-  \label{eq:decay_diff}
-\end{align}
-and, using \eqref{eq:san_bmu_mu},
-\begin{align}
-  \D\p{s_\mu} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
-  &= S_\mu^\ell \p{S - \ell - n - \mu S_\z} \p{1+\mu S_\z}^m S_\bmu^n
-  - S_\mu^\ell\sp{S - \f12\p{\ell+n} - \mu S_\z}
-  \p{\mu S_\z}^m S_\bmu^n \notag \\
-  &\qquad + \ell n \p{2S - \ell - n + 2}
-  S_\mu^{\ell-1} \p{\mu S_\z}^m S_\bmu^{n-1} \notag \\
-  &\qquad + \ell n \p{\ell-1} \p{n-1} S_\mu^{\ell-2} \p{S + \mu S_\z}
-  \p{-1+\mu S_\z}^m S_\bmu^{n-2}.
-  \label{eq:decay_same}
-\end{align}
-Decoherence via jump operators $s_\bmu^{(j)}$ only couples operators
-$S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$ to operators
-$S_\mu^\ell \p{\mu S_\z}^{m'} S_\bmu^n$ with $m'\le m$.  Decoherence
-via jump operators $s_\mu^{(j)}$, meanwhile, makes operators
-$S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$ ``grow'' in $m$ through the last
-term in \eqref{eq:decay_same}, although the sum $\ell+m+n$ does not
-grow.
-
-\subsection{Dephasing}
-\label{sec:dephasing_single}
-
-The effect of decoherence via single-spin dephasing is described by
-\begin{align}
-  \D\p{s_\z} \O
-  = \sum_j\p{s_\z^{(j)} \O s_\z^{(j)}
-    - \f12\sp{s_\z^{(j)} s_\z^{(j)},\O}_+}
-  = \sum_j s_\z^{(j)} \O s_\z^{(j)} - \f12 S \O.
-\end{align}
-From \eqref{eq:san_z_z}, we then have
-\begin{align}
-  \D\p{s_\z} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
-  = -\f12\p{\ell+n} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n
-  + \ell n S_\mu^{\ell-1} \p{S + \mu S_\z}
-  \p{-1 + \mu S_\z}^m S_\bmu^{n-1}.
-\end{align}
-Decoherence via single-spin dephasing makes operators
-$S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$ ``grow'' in $m$, although the sum
-$\ell+m+n$ does not grow.
-
-\subsection{The general case}
-\label{sec:general_single}
-
-The most general type of single-spin decoherence is described by
-\begin{align}
-  \D\p{g} \O
-  = \sum_j\p{{g^{(j)}}^\dag \O g^{(j)}
-    - \f12\sp{{g^{(j)}}^\dag g^{(j)}, \O}_+},
-  &&
-  g \equiv g_\z s_\z + g_+ s_+ + g_- s_-.
-  \label{eq:D_general_single}
-\end{align}
-To simplify \eqref{eq:D_general_single}, we expand
-\begin{align}
-  g^\dag \O g
-  = \abs{g_\z}^2 s_\z \O s_\z
-  + \sum_\mu \p{\abs{g_\mu}^2 s_\bmu \O s_\mu
-    + g_\bmu^* g_\mu s_\mu \O s_\mu
-    + g_\z^* g_\mu s_\z \O s_\mu
-    + g_\bmu^* g_\z s_\mu \O s_\z},
-\end{align}
-and
-\begin{align}
-  g^\dag g
-  = \f14 \abs{g_\z}^2
-  + \f12 \sum_\mu \sp{\abs{g_\mu}^2 \p{1-2\mu s_\z}
-    + \mu \p{g_\z^*g_\mu - g_\bmu^*g_\z} s_\mu},
-\end{align}
-which implies
-\begin{align}
-  \D\p{g} \O
-  &= \sum_{X\in\set{\z,+,-}} \abs{g_X}^2 \D\p{s_X} \O
-  + \sum_{\mu,j}
-  \p{g_\bmu^* g_\mu s_\mu^{(j)} \O s_\mu^{(j)}
-    + g_\z^* g_\mu s_\z^{(j)} \O s_\mu^{(j)}
-    + g_\bmu^* g_\z s_\mu^{(j)} \O s_\z^{(j)}}
-  \notag \\
-  &\qquad -\f14 \sum_\mu \mu
-  \p{g_\z^*g_\mu - g_\bmu^*g_\z} \sp{S_\mu, \O}_+.
-\end{align}
-In order to compute the effect of this decoherence on general
-collective spin operators, we expand the anti-commutator
-\begin{align}
-  \sp{S_\mu, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
-  = S_\mu^{\ell+1} \sp{\p{\mu S_\z}^m+\p{1+\mu S_\z}^m} S_\bmu^n
-  - n S_\mu^\ell \p{n-1+2\mu S_\z} \p{\mu S_\z}^m S_\bmu^{n-1}.
-  \label{eq:S_mu_acomm}
-\end{align}
-Recognizing a resemblance between terms in \eqref{eq:S_mu_acomm} and
-\eqref{eq:san_z_mu}, we collect terms to simplify
-\begin{align}
-  \sum_j s_\z^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\mu^{(j)}
-  - \f14 \mu \sp{S_\mu, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
-  = K_{\ell mn;\mu} + L_{\ell mn;\mu}
-  \label{eq:dec_z_mu}
-\end{align}
-and likewise
-\begin{align}
-  \sum_j s_\mu^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\z^{(j)}
-  + \f14 \mu \sp{S_\mu, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
-  = K_{\ell mn;\mu} + M_{\ell mn;\mu}
-  \label{eq:dec_mu_z}
-\end{align}
-with
-\begin{align}
-  K_{\ell mn;\mu}
-  &\equiv \f14 \mu S_\mu^{\ell+1}
-  \sp{\p{1+\mu S_\z}^m-\p{\mu S_\z}^m} S_\bmu^n, \\
-  L_{\ell mn;\mu}
-  &\equiv -\mu n S_\mu^\ell \sp{S-\ell-\f34\p{n-1}-\f12\mu S_\z}
-  \p{\mu S_\z}^m S_\bmu^{n-1}
-  - \mu\ell n\p{n-1} S_\mu^{\ell-1}
-  \p{S+\mu S_\z} \p{-1+\mu S_\z}^m S_\bmu^{n-2}, \\
-  M_{\ell mn;\mu}
-  &\equiv \mu n S_\mu^\ell \sp{\p{S+\mu S_\z}\p{-1+\mu S_\z}^m
-    - \f12\p{\f12\sp{n-1}+\mu S_\z}\p{\mu S_\z}^m} S_\bmu^{n-1}.
-\end{align}
-Defining for completion
-\begin{align}
-  P_{\ell mn;\mu}
-  &\equiv \sum_j s_\mu^{(j)} S_\mu^\ell
-  \p{\mu S_\z}^m S_\bmu^n s_\mu^{(j)}
-  = n S_\mu^{\ell+1} \p{\mu S_\z}^m S_\bmu^{n-1}
-  - n\p{n-1} S_\mu^\ell \p{S+\mu S_\z} \p{-1+\mu S_\z}^m S_\bmu^{n-2},
-\end{align}
-and
-\begin{align}
-  Q_{\ell mn;\mu}^{(g)}
-  \equiv g_\bmu^* g_\mu P_{\ell mn;\mu}
-  + \p{g_\z^* g_\mu + g_\bmu^* g_\z}
-  K_{\ell mn;\mu}
-  + g_\z^* g_\mu L_{\ell mn;\mu}
-  + g_\bmu^* g_\z M_{\ell mn;\mu},
-  \label{eq:Q_single}
-\end{align}
-we finally have
-\begin{align}
-  \D\p{g} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
-  = \sum_{X\in\set{\z,+,-}} \abs{g_X}^2
-  \D\p{s_X} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
-  + Q_{\ell mn;\mu}^{(g)} + \sp{Q_{nm\ell;\mu}^{(g)}}^\dag.
-\end{align}
-Note that the sum $\ell+m+n$ for operators
-$S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$ does not grow under this type of
-decoherence.
-
-\section{Sandwich identities for collective-spin decoherence
-  calculations}
-\label{sec:sandwich_collective}
-
-In analogy with the work in Appendix \ref{sec:sandwich_single}, in
-this section we work out sandwich identities necessary for
-collective-spin decoherence calculations.  The simplest cases are
-\begin{align}
-  S_\mu S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\bmu
-  &= S_\mu^{\ell+1} \p{\mu S_\z}^m S_\bmu^{n+1},
-  \allowdisplaybreaks \\
-  S_\mu S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\z
-  &= \mu S_\mu^{\ell+1} \p{n+\mu S_\z} \p{\mu S_\z}^m S_\bmu^n,
-  \allowdisplaybreaks \\
-  S_\z S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\z
-  &= S_\mu^\ell \sp{\ell n + \p{\ell+n} \mu S_\z + \p{\mu S_\z}^2}
-  \p{\mu S_\z}^m S_\bmu^n.
-\end{align}
-With a bit more work, we can also find
-\begin{align}
-  S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\mu
-  &= S_\mu^{\ell+1} \p{1+\mu S_\z}^m S_\bmu^n
-  - n S_\mu^\ell \p{n-1+2\mu S_\z} \p{\mu S_\z}^m S_\bmu^{n-1},
-\end{align}
-which implies
-\begin{align}
-  S_\mu S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\mu
-  &= S_\mu^{\ell+2} \p{1+\mu S_\z}^m S_\bmu^n
-  - n S_\mu^{\ell+1} \p{n-1+2\mu S_\z} \p{\mu S_\z}^m S_\bmu^{n-1},
-  \allowdisplaybreaks \\
-  S_\z S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\mu
-  &= \mu S_\mu^{\ell+1} \p{\ell+1+\mu S_\z} \p{1+\mu S_\z}^m S_\bmu^n
-  \notag \\
-  &\qquad - \mu n S_\mu^\ell
-  \sp{\ell\p{n-1} + \p{2\ell+n-1}\mu S_\z + 2\p{\mu S_\z}^2}
-  \p{\mu S_\z}^m S_\bmu^{n-1}.
-\end{align}
-Finally, we compute
-\begin{align}
-  S_\bmu S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\mu
-  &= \sp{S_\mu^\ell S_\bmu - \ell S_\mu^{\ell-1} \p{\ell-1+2\mu S_\z}}
-  \p{\mu S_\z}^m
-  \sp{S_\mu S_\bmu^n - n \p{n-1+2\mu S_\z} S_\bmu^{n-1}} \notag \\
-  &= S_\mu^\ell S_\bmu \p{\mu S_\z}^m S_\mu S_\bmu^n \notag \\
-  &\qquad - S_\mu^\ell
-  \sp{\ell\p{\ell+1} + n\p{n+1}+2\p{\ell+n}\mu S_\z}
-  \p{1+\mu S_\z}^m S_\bmu^n \notag \\
-  &\qquad + \ell n S_\mu^{\ell-1}
-  \sp{\p{\ell-1}\p{n-1}+2\p{\ell+n-2}\mu S_\z + 4\p{\mu S_\z}^2}
-  \p{\mu S_\z}^m S_\bmu^{n-1},
-\end{align}
-where
-\begin{multline}
-  S_\bmu \p{\mu S_\z}^m S_\mu
-  = S_\bmu S_\mu \p{1+\mu S_\z}^m
-  = \p{S_\mu S_\bmu - 2\mu S_\z} \p{1+\mu S_\z}^m
-  = S_\mu \p{2+\mu S_\z}^m S_\bmu - 2\mu S_\z \p{1+\mu S_\z}^m,
-\end{multline}
-so
-\begin{align}
-  S_\bmu S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\mu
-  &= S_\mu^{\ell+1} \p{2+\mu S_\z}^m S_\bmu^{n+1} \notag \\
-  &\qquad - S_\mu^\ell
-  \sp{\ell\p{\ell+1} + n\p{n+1}+2\p{\ell+n+1}\mu S_\z}
-  \p{1+\mu S_\z}^m S_\bmu^n \notag \\
-  &\qquad + \ell n S_\mu^{\ell-1}
-  \sp{\p{\ell-1}\p{n-1}+2\p{\ell+n-2}\mu S_\z + 4\p{\mu S_\z}^2}
-  \p{\mu S_\z}^m S_\bmu^{n-1}.
-\end{align}
-
-\section{Collective spin decoherence}
-\label{sec:decoherence_collective}
-
-In this section we work out the effects of collective decoherence on
-general collective spin operators.  For shorthand, we define
-\begin{align}
-  \D\p{G} \O
-  \equiv \D\p{\set{G}} \O
-  = G^\dag \O G - \f12\sp{G^\dag G, \O}_+,
-\end{align}
-where $G$ is a collective spin jump operator.
-
-\subsection{Decay-type decoherence and dephasing}
-\label{sec:decay_dephasing_collective}
-
-Making use of the results in Appendix \ref{sec:sandwich_collective},
-we find that the effects of collective decay-type decoherence on
-general collective spin operators are given by
-\begin{align}
-  \D\p{S_\bmu} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
-  &= -S_\mu^{\ell+1} \sp{\p{1+\mu S_\z}^m - \p{\mu S_\z}^m} S_\bmu^{n+1}
-  \notag \\
-  &\qquad + \f12 S_\mu^\ell \sp{\ell\p{\ell-1} + n\p{n-1}
-    + 2\p{\ell+n}\mu S_\z} \p{\mu S_\z}^m S_\bmu^n,
-\end{align}
-and
-\begin{align}
-  \D\p{S_\mu} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
-  &= S_\mu^{\ell+1} \sp{\p{2+\mu S_\z}^m-\p{1+\mu S_\z}^m} S_\bmu^{n+1}
-  \notag \\
-  &\qquad - S_\mu^\ell
-  \sp{\ell\p{\ell+1} + n\p{n+1}+2\p{\ell+n+1}\mu S_\z}
-  \p{1+\mu S_\z}^m S_\bmu^n \notag \\
-  &\qquad + \f12 S_\mu^\ell
-  \sp{\ell\p{\ell+1} + n\p{n+1}+2\p{\ell+n+2}\mu S_\z}
-  \p{\mu S_\z}^m S_\bmu^n \notag \\
-  &\qquad + \ell n S_\mu^{\ell-1}
-  \sp{\p{\ell-1}\p{n-1}+2\p{\ell+n-2}\mu S_\z + 4\p{\mu S_\z}^2}
-  \p{\mu S_\z}^m S_\bmu^{n-1}.
-\end{align}
-Similarly, the effect of collective dephasing is given by
-\begin{align}
-  \D\p{S_\z} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
-  = -\f12 \p{\ell-n}^2 S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n.
-\end{align}
-
-\subsection{The general case}
-\label{sec:general_collective}
-
-More generally, we consider jump operators of the form
-\begin{align}
-  G \equiv G_\z S_\z + G_+ S_+ + G_- S_-,
-\end{align}
-whose decoherence effects are determined by
-\begin{align}
-  G^\dag \O G
-  = \abs{G_\z}^2 S_\z \O S_\z
-  + \sum_\mu \p{\abs{G_\mu}^2 S_\bmu \O S_\mu
-    + G_\bmu^* G_\mu S_\mu \O S_\mu
-    + G_\z^* G_\mu S_\z \O S_\mu
-    + G_\bmu^* G_\z S_\mu \O S_\z},
-\end{align}
-and
-\begin{align}
-  G^\dag G
-  = \abs{G_\z}^2 S_\z^2
-  + \sum_\mu \p{\abs{G_\mu}^2 S_\bmu S_\mu
-    + G_\z^*G_\mu S_\z S_\mu
-    + G_\bmu^* G_\z S_\mu S_\z
-    + G_\bmu^* G_\mu S_\mu^2},
-\end{align}
-which implies
-\begin{align}
-  \D\p{G} \O
-  &= \sum_{X\in\set{\z,+,-}} \abs{G_X}^2 \D\p{S_X} \O
-  + \sum_\mu \p{G_\bmu^* G_\mu S_\mu \O S_\mu
-    + G_\z^* G_\mu S_\z \O S_\mu
-    + G_\bmu^* G_\z S_\mu \O S_\z}
-  \notag \\
-  &\qquad -\f12 \sum_\mu\p{G_\bmu^* G_\mu \sp{S_\mu^2, \O}_+
-    + G_\z^*G_\mu \sp{S_\z S_\mu, \O}_+
-    + G_\bmu^* G_\z \sp{S_\mu S_\z, \O}_+}.
-\end{align}
-In order to compute the effect of this decoherence on general
-collective spin operators, we expand the anti-commutators
-\begin{align}
-  \sp{S_\mu^2, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
-  &= S_\mu^{\ell+2} \sp{\p{2+\mu S_\z}^m+\p{\mu S_\z}^m} S_\bmu^n
-  - 2n S_\mu^{\ell+1} \p{n+2\mu S_\z} \p{1+\mu S_\z}^m S_\bmu^{n-1}
-  \notag \\
-  &\qquad + n\p{n-1} S_\mu^\ell \sp{\p{n-1}\p{n-2}
-    + 2\p{2n-3}\mu S_\z + 4\p{\mu S_\z}^2} \p{\mu S_\z}^m S_\bmu^{n-2},
-  \allowdisplaybreaks \\
-  \sp{S_\z S_\mu, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
-  &= \mu S_\mu^{\ell+1} \sp{\p{\ell+1+\mu S_\z}\p{\mu S_\z}^m
-    + \p{n+1+\mu S_\z} \p{1+\mu S_\z}^m } S_\bmu^n \notag \\
-  &\qquad - \mu n S_\mu^\ell \sp{n \p{n-1}
-    + \p{3n-1}\mu S_\z + 2\p{\mu S_\z}^2} \p{\mu S_\z}^m S_\bmu^{n-1},
-  \allowdisplaybreaks \\
-  \sp{S_\mu S_\z, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
-  &= \mu S_\mu^{\ell+1} \sp{\p{\ell+\mu S_\z}\p{\mu S_\z}^m
-    + \p{n+\mu S_\z} \p{1+\mu S_\z}^m} S_\bmu^n \notag \\
-  &\qquad - \mu n S_\mu^\ell \sp{\p{n-1}^2
-    + 3\p{n-1}\mu S_\z + 2\p{\mu S_\z}^2} \p{\mu S_\z}^m S_\bmu^{n-1}.
-\end{align}
-Collecting terms and defining
-\begin{align}
-  G_{\z,\mu}^{(\pm)}
-  &\equiv \f12\p{G_\z^* G_\mu \pm G_\bmu^* G_\z},
-  \allowdisplaybreaks \\
-  \tilde L_{\ell mn;\mu}^{(G)}
-  &\equiv \mu \sp{\p{\ell-n+\f12} G_{\z,\mu}^{(+)}
-    + \p{\ell+\f12} G_{\z,\mu}^{(-)}}
-  S_\mu^{\ell+1} \p{1+\mu S_\z}^m S_\bmu^n \notag \\
-  &\qquad -\mu \sp{\p{\ell-n+\f12} G_{\z,\mu}^{(+)}
-    + \p{n+\f12} G_{\z,\mu}^{(-)}}
-  S_\mu^{\ell+1} \p{\mu S_\z}^m S_\bmu^n \notag \\
-  &\qquad + \mu G_{\z,\mu}^{(-)}
-  S_\mu^{\ell+1} \mu S_\z
-  \sp{\p{1+\mu S_\z}^m - \p{\mu S_\z}^m} S_\bmu^n,
-  \allowdisplaybreaks \\
-  \tilde M_{\ell mn;\mu}^{(G)}
-  &= -\mu n\p{n-1} \sp{\p{\ell-n+\f12} G_{\z,\mu}^{(+)}
-    + \p{\ell-\f12} G_{\z,\mu}^{(-)}}
-  S_\mu^\ell \p{\mu S_\z}^m S_\bmu^{n-1} \notag \\
-  &\qquad - 2\mu n \sp{\p{\ell-n+\f12} G_{\z,\mu}^{(+)}
-    + \p{\ell+\f12n-1} G_{\z,\mu}^{(-)}}
-  S_\mu^\ell \p{\mu S_\z}^{m+1} S_\bmu^{n-1} \notag \\
-  &\qquad - 2\mu n G_{\z,\mu}^{(-)}
-  S_\mu^\ell \p{\mu S_\z}^{m+2} S_\bmu^{n-1},
-  \allowdisplaybreaks \\
-  \tilde P_{\ell mn;\mu}
-  &\equiv -\f12 S_\mu^{\ell+2}
-  \sp{\p{2+\mu S_\z}^m - 2\p{1+\mu S_\z}^m + \p{\mu S_\z}^m}
-  S_\bmu^n \notag \\
-  &\qquad + n S_\mu^{\ell+1} \sp{\p{n+2\mu S_\z} \p{1+\mu S_\z}^m
-    - \p{n-1+2\mu S_\z} \p{\mu S_\z}^m}
-  S_\bmu^{n-1} \notag \\
-  &\qquad -n\p{n-1} S_\mu^\ell
-  \sp{\f12\p{n-1}\p{n-2} + \p{2n-3}\mu S_\z + 2\p{\mu S_\z}^2}
-  \p{\mu S_\z}^m S_\bmu^{n-2},
-  \allowdisplaybreaks \\
-  \tilde Q_{\ell mn;\mu}^{(G)}
-  &\equiv G_\bmu^* G_\mu \tilde P_{\ell mn;\mu}
-  + \tilde L_{\ell mn;\mu}^{(G)}
-  + \tilde M_{\ell mn;\mu}^{(G)},
-\end{align}
-we then have
-\begin{align}
-  \D\p{G} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
-  = \sum_{X\in\set{\z,+,-}} \abs{G_X}^2
-  \D\p{S_X} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
-  + \tilde Q_{\ell mn;\mu}^{(G)}
-  + \sp{\tilde Q_{nm\ell;\mu}^{(G)}}^\dag.
-\end{align}
-Note that the sum $\ell+m+n$ for operators
-$S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$ grows by one if $G_\mu\ne0$ or
-$G_\bmu\ne0$, and does not grow otherwise.
-
-\section{Initial conditions}
-\label{sec:initial_conditions}
-
-Here we compute the expectation values of collective spin operators
-with respect to spin-polarized (also Gaussian, or spin-coherent)
-states.  These states are parameterized by polar and azimuthal angles
-$\theta\in[0,\pi)$, $\phi\in[0,2\pi)$, and lie within the Dicke
-manifold spanned by states
-$\ket{k}\propto S_+^{S+k}\ket{\dn}^{\otimes N}$ with $S\equiv N/2$ and
-$S_\z\ket{k}=k\ket{k}$:
-\begin{align}
-  \ket{\theta,\phi}
-  \equiv \sp{\cos\p{\theta/2} e^{-i\phi/2} \ket\up
-    + \sin\p{\theta/2} e^{i\phi/2} \ket\dn}^{\otimes N}
-  = \sum_{k=-S}^S { N \choose S+k }^{1/2}
-  \cos\p{\theta/2}^{S+k} \sin\p{\theta/2}^{S-k} e^{-ik\phi} \ket{k}.
-\end{align}
-We can likewise expand, within the Dicke manifold,
-\begin{align}
-  S_\z = \sum_{k=-S}^S k \op{k},
-  &&
-  S_\mu = \sum_{k=-S+\delta_{\mu,-1}}^{S-\delta_{\mu,1}}
-  g_\mu\p{k} \op{k+\mu}{k}
-  = \sum_{k=-S+\delta_{\bmu,-1}}^{S-\delta_{\bmu,1}}
-  g_\bmu\p{k} \op{k}{k+\bmu},
-\end{align}
-where $\bmu\equiv-\mu\in\set{+1,-1}$ and
-\begin{align}
-  g_\mu\p{k} \equiv \sqrt{\p{S-\mu k}\p{S+\mu k+1}},
-\end{align}
-which implies
-\begin{align}
-  S_\mu^\ell \p{\mu S_\z^m} S_\bmu^n
-  &= \sum_{k=-S+\delta_{\mu,-1}\max\set{\ell,n}}
-  ^{S-\delta_{\mu,1}\max\set{\ell,n}} \p{\mu k}^m
-  \sp{\prod_{p=0}^{\ell-1} g_\mu\p{k+\mu p}}
-  \sp{\prod_{q=0}^{n-1} g_\mu\p{k+\mu q}}
-  \op{k+\mu\ell}{k+\mu n} \\
-  &= \sum_{\mu k=-\mu S-\delta_{\mu,-1}\max\set{\ell,n}}
-  ^{\mu S-\delta_{\mu,1}\max\set{\ell,n}} \p{\mu k}^m
-  \f{\p{S-\mu k}!}{\p{S+\mu k}!}
-  \sp{\f{\p{S+\mu k+\ell}!}{\p{S-\mu k-\ell}!}
-    \f{\p{S+\mu k+n}!}{\p{S-\mu k-n}!}}^{1/2}
-  \op{k+\mu\ell}{k+\mu n} \\
-  &= \sum_{k=-S}^{S-\max\set{\ell,n}} k^m
-  \f{\p{S-k}!}{\p{S+k}!}
-  \sp{\f{\p{S+k+\ell}!}{\p{S-k-\ell}!}
-    \f{\p{S+k+n}!}{\p{S-k-n}!}}^{1/2}
-  \op{\mu\p{k+\ell}}{\mu\p{k+n}}.
-\end{align}
-This expansion allows us to compute the expectation value
-\begin{align}
-  \bk{\theta,\phi|S_\mu^\ell \p{\mu S_\z^m} S_\bmu^n|\theta,\phi}
-  &= e^{i\phi \mu\p{\ell-n}} N! \sum_{k=-S}^{S-\max\set{\ell,n}}
-  \f{k^m \p{S-k}! f_{\mu\ell n}\p{k,\theta}}
-  {\p{S+k}! \p{S-k-\ell}! \p{S-k-n}!} \\
-  &= e^{i\phi \mu\p{\ell-n}} \p{-1}^m N! \sum_{k=0}^{N-\max\set{\ell,n}}
-  \f{\p{S-k}^m \p{N-k}! \tilde f_{\mu\ell n}\p{k,\theta}}
-  {k! \p{N-k-\ell}! \p{N-k-n}!}
-\end{align}
-where
-\begin{align}
-  f_{\mu\ell n}\p{k,\theta}
-  \equiv \cos\p{\theta/2}^{N+\mu\p{2k+\ell+n}}
-  \sin\p{\theta/2}^{N-\mu\p{2k+\ell+n}},
-\end{align}
-\begin{align}
-  \tilde f_{\mu\ell n}\p{k,\theta}
-  \equiv f_{\mu\ell n}\p{k-S,\theta}
-  = \cos\p{\theta/2}^{2N\delta_{\mu,-1}+\mu\p{2k+\ell+n}}
-  \sin\p{\theta/2}^{2N\delta_{\mu,1}-\mu\p{2k+\ell+n}}.
-\end{align}
-Defining the states
-\begin{align}
-  \ket{+\Z} \equiv \ket{0,0} = \ket\up^{\otimes N}, &&
-  \ket{-\Z} \equiv \ket{\pi,0} = \ket\dn^{\otimes N}, &&
-  \ket\X \equiv \ket{\pi/2,0}
-  = \p{\f{\ket\up+\ket\dn}{\sqrt2}}^{\otimes N},
-\end{align}
-some particular expectation values of interest are
-\begin{align}
-  \bk{\nu\Z|S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n|\nu\Z}
-  = \delta_{\ell n} \times
-  \begin{cases}
-    \p{S-n}^m \f{N! n!}{\p{N-n}!} & \mu = \nu, \\
-    \delta_{n,0} \p{-S}^m & \mu \ne \nu,
-  \end{cases},
-\end{align}
-and
-\begin{align}
-  \bk{\X|S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n|\X}
-  = \p{-1}^m \f{N!}{2^N} \sum_{k=0}^{N-\max\set{\ell,n}}
-  \f{\p{S-k}^m \p{N-k}!}{k!\p{N-k-\ell}!\p{N-k-n}!}.
-\end{align}
-
-\section{Computing correlators with the truncated short-time (TST)
-  expansion}
-\label{sec:tutorial}
-
-Here, we provide a pedagogical tutorial for computing correlators
-using the truncated short-time TST expansion.  For concreteness, we
-nominally consider $N$ spins evolving under the one-axis twisting
-(OAT) Hamiltonian
-\begin{align}
-  H_{\t{OAT}} = \chi S_\z^2,
-\end{align}
-additionally subject to spontaneous single-spin decay at rate
-$\gamma_-$, with jump operators $\J_-=\set{s_-^{(j)}:j=1,2,\cdots,N}$.
-The equation of motion for a Heisenberg operator
-$S_+^\ell S_\z^m S_-^n$ is
-\begin{align}
-  \f{d}{dt} \bk{S_+^\ell S_\z^m S_-^n}
-  = i\chi\Bk{\sp{S_\z^2, S_+^\ell S_\z^m S_-^n}_-}
-  + \gamma_- \Bk{\D\p{\J_-} \p{S_+^\ell S_\z^m S_-^n}},
-  \label{eq:tutorial_EOM}
-\end{align}
-where we can use the results in appendices \ref{sec:general_product}
-and \ref{sec:decoherence_single} respectively to evaluate the
-commutator $\sp{S_\z^2, S_+^\ell S_\z^m S_-^n}_-$ and dissipator
-$\D\p{\J_-} \p{S_+^\ell S_\z^m S_-^n}$ in \eqref{eq:tutorial_EOM} (see
-\eqref{eq:general_product} and \eqref{eq:decay_diff} in particular),
-giving us
-\begin{multline}
-  \f{d}{dt} \bk{S_+^\ell S_\z^m S_-^n} \\
-  = i\chi \bk{\p{\ell-n} S_+^\ell \p{\ell+n+2S_\z} S_\z^m S_-^n}
-  + \gamma_- \Bk{S_+^\ell \sp{\p{S+S_\z}\p{-1+S_\z}^m
-      - \p{S+\f{\ell+n}{2}+S_\z} S_\z^m} S_-^n}.
-  \label{eq:tutorial_EOM_expanded}
-\end{multline}
-In practice, we do not want to keep track of such an expansion by
-hand, especially in the case of e.g.~the two-axis twisting (TAT) and
-twist-and-turn (TNT) models with more general types of decoherence,
-for which the analogue of \eqref{eq:tutorial_EOM_expanded} may take
-several lines just to write out in full.  Defining the operators
-$\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$ with
-$\v m\equiv\p{m_+,m_\z,m_-}$ for shorthand, we note that the vector
-space spanned by $\set{\S_{\v m}}$ is closed under time evolution.  We
-therefore expand
-\begin{align}
-  \f{d}{dt} \bk{\S_{\v n}}
-  = \bk{T \S_{\v n}}
-  = \sum_{\v m} \bk{\S_{\v m}} T_{\v m\v n},
-  \label{eq:tutorial_EOM_general}
-\end{align}
-where $T$ is a superoperator that generates time evolution for
-Heisenberg operators.  In the present example, the matrix elements
-$T_{\v m\v n}\in\C$ of $T$ are defined by
-\eqref{eq:tutorial_EOM_expanded} and \eqref{eq:tutorial_EOM_general}.
-For any Hamiltonian $H$ with decoherence characterized by sets of jump
-operators $\J$ and decoherence rates $\gamma_\J$, the matrix elements
-$T_{\v m\v n}$ are more generally defined by
-\begin{align}
-  \bk{T \S_{\v n}}
-  = i\bk{\sp{H,\S_{\v n}}_-}
-  + \sum_\J \gamma_\J \bk{\D\p{\J} \S_{\v n}}
-  = \sum_{\v m} \bk{\S_{\v m}} T_{\v m\v n}.
-\end{align}
-The results in Appendices \ref{sec:general_product},
-\ref{sec:decoherence_single}, and \ref{sec:decoherence_collective} can
-be used to write model-agnostic codes that compute matrix elements
-$T_{\v m\v n}$, taking a particular Hamiltonian $H$ and decoherence
-processes $\set{\p{\J,\gamma_\J}}$ as inputs.
-
-In order to compute a quantity such as spin squeezing, we need to
-compute correlators of the form $\bk{\S_{\v n}\p{t}}$, where
-$\S_{\v n}\p{t}$ is the time-evolved Heisenberg operator corresponding
-to $\S_{\v n}$, with $\S_{\v n}\p{0}\equiv\S_{\v n}$.  The order-$M$
-truncated short-time (TST) expansion takes
-\begin{align}
-  \bk{\S_{\v n}\p{t}}
-  = \sum_{k\ge0} \f{t^k}{k!} \bk{T^k\S_{\v n}\p{0}}
-  = \sum_{k\ge0} \f{t^k}{k!}
-  \sum_{\v m} \bk{\S_{\v m}\p{0}} T^k_{\v m\v n}
-  \to \sum_{k=0}^M \f{t^k}{k!}
-  \sum_{\v m} \bk{\S_{\v m}\p{0}} T^k_{\v m\v n},
-  \label{eq:tutorial_TST}
-\end{align}
-where $T^k_{\v m\v n}$ are matrix elements of the $k$-th time
-derivative operator $T^k$, given by
-\begin{align}
-  T^0_{\v m\v n} \equiv
-    \begin{cases}
-      1 & \v m = \v n, \\
-      0 & \t{otherwise}
-    \end{cases},
-  &&
-  T^1_{\v m\v n} \equiv T_{\v m\v n},
-  &&
-  T^{k>1}_{\v m\v n}
-  \equiv \sum_{\v p_1,\v p_2,\cdots,\v p_{k-1}}
-  T_{\v m\v p_{k-1}} \cdots T_{\v p_3\v p_2}
-  T_{\v p_2\v p_1} T_{\v p_1\v n}.
-\end{align}
-Matrix elements $T^k_{\v m\v n}$ and initial-time expectation values
-$\bk{\S_{\v m}\p{0}}$ are thus computed as needed for any particular
-correlator $\bk{\S_{\v n}\p{t}}$ of interest, and combined according
-to \eqref{eq:tutorial_TST}.  Note that initial-time expectation values
-$\bk{\S_{\v m}\p{0}}$ are an input to the TST expansion, and need to
-be computed separately for any given initial state; expectation values
-with respect to spin-polarized (Gaussian) states are provided in
-Appendix \ref{sec:initial_conditions}.  In practice, we further
-collect terms in \eqref{eq:tutorial_TST} to write
-\begin{align}
-  \bk{\S_{\v n}\p{t}} \to \sum_{k=0}^M c_{\v n k} t^k,
-  &&
-  c_{\v n k}
-  \equiv \f1{k!} \sum_{\v m} \bk{\S_{\v m}\p{0}} T^k_{\v m\v n},
-  \label{eq:tutorial_corr}
-\end{align}
-where $c_{\v n k}$ are time-independent coefficients for the
-polynomial time expansion of $\bk{\S_{\v n}\p{t}}$.  After computing
-the coefficients $c_{\v n k}$, there is only negligible computational
-overhead to compute the correlator $\bk{\S_{\v n}\p{t}}$ for any time
-$t$ up to the breakdown time $t_M$ of an order-$M$ TST expansion.
-
-\section{Short-time linear response and two-time correlators}
-\label{sec:linear_response}
-
-Here, we discuss the appearance of two-time correlation functions in
-the short-time linear response of correlators to perturbations of a
-Hamiltonian.  Consider an initial Hamiltonian $H$ perturbed by an
-operator $V$ with $\norm{V}\ll\norm{H}$, where $\norm{\O}$ denotes the
-operator norm of $\O$, such that the net Hamiltonian is
-$\widetilde H=H+V$.  We denote the generator of Heisenberg time
-evolution under the perturbed (unperturbed) Hamiltonian by
-$\widetilde T$ ($T$).  These generators are related by
-\begin{align}
-  \widetilde T = T + i\check V
-\end{align}
-where $\check V$ is a superoperator whose action on operators $\O$ is
-defined by
-\begin{align}
-  \check V \O \equiv \sp{V,\O}_-.
-\end{align}
-Through quadratic order in the time $t$ and linear order in the
-perturbation $\check V$, we can say that
-\begin{align}
-  e^{t\widetilde T}
-  \approx \f12\sp{e^{tT}, e^{it\check V}}_+
-  \approx e^{tT} + \f12 it \sp{e^{t T}, \check V}_+.
-\end{align}
-Defining perturbed and unperturbed Heisenberg operators
-$\widetilde\O\p{t}$ and $\O\p{t}$ that respectively satisfy
-$\bk{\widetilde\O\p{t}}=\bk{e^{t\widetilde T}\O}$ and
-$\bk{\O\p{t}}=\bk{e^{tT}\O}$, we thus find that for sufficiently small
-times $t$ and weak perturbations $V$,
-\begin{align}
-  \bk{\widetilde\O\p{t} - \O\p{t}}
-  = \Bk{\p{e^{t\widetilde T} - e^{tT}} \O}
-  \approx \f12 i t \p{\bk{\sp{V,\O}_-\p{t}}
-    + \bk{\sp{V,\O\p{t}}_-}}.
-  \label{eq:response}
-\end{align}
-Two-time correlators $\bk{V\O\p{t}}$ and $\bk{\O\p{t}V}$, in addition
-to the expectation values $\bk{\p{V\O}\p{t}}$ and
-$\bk{\p{\O V}\p{t}}$, thus determine the short-time linear response of
-correlators $\bk{\O\p{t}}$ to perturbations $V$ of a Hamiltonian.
-
-% \section{Changing operator bases, and a general product for algebras
-%   of permutationally-symmetric operators}
-% \label{sec:new_basis}
-
-% For a collective spin system composed of many small spins, the
-% operators $\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$ contain
-% $\abs{\v m}$-body operators (for $\abs{\v m}\equiv\sum_\mu m_\mu$),
-% but also carry lots of ``baggage'' in the form of $k$-body operators
-% with $k<\abs{\v m}$.  To cut down on this overhead, we try to change
-% our basis for collective spin operators from $\S_{\v m}$ to the
-% purely-$\abs{\v m}$-body operators
-% \begin{align}
-%   \tilde\S_{\v m} \equiv \sum_{\p{\v j;\v m}} P_{\v j},
-%   &&
-%   P_{\v j} \equiv \prod_{\mu,a} s_\mu^{\p{j^\mu_a}},
-%   \label{eq:S_P}
-% \end{align}
-% where $\v m \equiv \p{m_+,m_\z,m_-}\in\mathbb{N}_0^3$ with $m_\mu$ the
-% number of $s_\mu$ operators in each term of $\S_{\v m}$; $\v j$ is an
-% ordered list of indices for the spins addressed by $P_{\v j}$; and the
-% sum over $\p{\v j;\v m}$ denotes a sum over all possible $\v j$ for
-% which all indices in $\v j$ are distinct and
-% $\abs{\set{j^\mu_a}} = m_\mu$, such that $j^\mu_a$ indexes the $a$-th
-% spin addressed by an $s_\mu$ operator in $P_{\v j}$.
-
-% In order to make use of the operators $\tilde S_{\v m}$, we will need
-% to compute products of the form $\tilde S_{\v m} \tilde S_{\v n}$.  We
-% therefore expand such a product into terms that have exactly $s$ spins
-% addressed by both operators:
-% \begin{align}
-%   \tilde\S_{\v m} \tilde\S_{\v n}
-%   = \sum_{s\ge0} \sum_{\p{\v j,\v k;\v m,\v n,s}} P_{\v j} P_{\v k},
-%   &&
-%   \sum_{\p{\v j,\v k;\v m,\v n,s}} X \equiv
-%   \sum_{\substack{\p{\v j;\v m},\p{\v k;\v n} \\
-%       \abs{\set{j_\a^\mu}\cap\set{k_\b^\nu}}=s}} X.
-%   \label{eq:SS_PP}
-% \end{align}
-% Collecting terms in which $r_{\mu\nu}$ of the $s_\mu$ operators in
-% $P_{\v j}$ address the same spin as an $s_\nu$ operator in $P_{\v k}$,
-% we have
-% \begin{align}
-%   \tilde\S_{\v m} \tilde\S_{\v n}
-%   = \sum_{s\ge0} \sum_{\p{\v r;\v m,\v n,s}} f_{\v m\v n\v r}
-%   \sum_{\p{\v J,\v K;\v m,\v n,\v r}} P_{\v J} Q_{\v K},
-%   \label{eq:SS_PQ}
-% \end{align}
-% where the sum over $\p{\v r;\v m,\v n,s}$ denotes a sum over all
-% values of $\v r$ with the restrictions
-% \begin{align}
-%   r_{\mu\nu} \ge 0 ~\forall~ \mu,\nu,
-%   &&
-%   \sum_\nu r_{\mu\nu} \le m_\mu,
-%   &&
-%   \sum_\mu r_{\mu\nu} \le n_\nu,
-%   &&
-%   \sum_{\mu,\nu} r_{\mu\nu} = s;
-%   \label{eq:rest_r}
-% \end{align}
-% the factor $f_{\v m\v n\v r}$ counts the number of ways to pair
-% operators in $P_{\v j}$ and $P_{\v k}$ according to $\v r$, or
-% \begin{align}
-%   f_{\v m\v n\v r}
-%   &\equiv \sp{\prod_{\mu,\nu}
-%     { m_\mu - \sum_{\rho<\nu} r_{\mu\rho} \choose r_{\mu\nu} }
-%     { n_\mu - \sum_{\rho<\nu} r_{\rho\mu} \choose r_{\nu\mu} }}
-%   \sp{\prod_{\mu,\nu} r_{\mu\nu}!} \\
-%   &= \sp{\prod_\mu \f{m_\mu!}{\p{m_\mu-\sum_\rho r_{\mu\rho}}!}
-%     \f{n_\mu!}{\p{n_\mu-\sum_\rho r_{\rho\mu}}!}}
-%   \sp{\prod_{\mu,\nu} r_{\mu\nu}!}^{-1};
-% \end{align}
-% $\v J$ and $\v K$ are respectively ordered lists of indices for spins
-% addressed by one and two single-spin operators in a term of the
-% product $P_{\v j} P_{\v k}$ with fixed $\v r$; the sum over
-% $\p{\v J,\v K;\v m,\v n,\v r}$ denotes a sum over all values of
-% $\v J,\v K$ with the restrictions
-% \begin{align}
-%   \abs{\set{J_a^\mu}}
-%   = m_\mu + n_\mu - \sum_\rho\p{r_{\mu\rho}+r_{\rho\mu}},
-%   &&
-%   \abs{\set{K^{\mu\nu}_a}} = r_{\mu\nu},
-%   &&
-%   \set{J_a^\lambda} \cap \set{K_b^{\mu\nu}} = \varnothing,
-% \end{align}
-% with $\varnothing$ denoting the empty set; and finally, similarly to
-% $P_{\v j}$ in \eqref{eq:S_P} we define
-% \begin{align}
-%   Q_{\v K} \equiv \prod_{\mu,\nu,a}
-%   s_\mu^{(K^{\mu\nu}_a)} s_\nu^{(K^{\mu\nu}_a)}.
-% \end{align}
-% As the sum in \eqref{eq:SS_PQ} is invariant under permutation of the
-% indices in $\v K$, we can safely neglect keeping track of these
-% indices and simply write
-% \begin{align}
-%   Q_{\v K}
-%   = \prod_{\mu,\nu} \prod_{a=1}^{r_{\mu\nu}}
-%   s_\mu^{(K^{\mu\nu}_a)} s_\nu^{(K^{\mu\nu}_a)}
-%   \simeq \bigotimes_{\mu,\nu} \bigotimes_{a=1}^{r_{\mu\nu}} s_\mu s_\nu
-%   = \bigotimes_{\mu,\nu} \bigotimes_{a=1}^{r_{\mu\nu}}
-%   \sum_\rho \eta_{\mu\nu\rho} s_\rho,
-%   \label{eq:Q_K_eta}
-% \end{align}
-% where $\simeq$ denotes equality up to a re-indexing of spins and
-% additional tensor factors of the single-spin identity operator
-% (i.e.~$\1$); we have introduced explicit dependence on
-% $r_{\mu\nu}=\abs{\set{K^{\mu\nu}_a}}$ for brevity; and
-% $\eta_{\mu\nu\rho}$ is a structure constant defined by
-% $s_\mu s_\nu=\sum_\rho\eta_{\mu\nu\rho}s_\rho$.  Unlike the sums over
-% $\mu,\nu$ in most of the work above, the sum over $\rho$ here includes
-% an index for the identity operator $s_0\equiv\1$.  Distributing the
-% product of sums into a sum of products gives us
-% \begin{align}
-%   Q_{\v K}
-%   \simeq \sum_{\v\rho} \bigotimes_{\mu,\nu,a}
-%   \eta_{\mu\nu\rho^{\mu\nu}_a} s_{\rho^{\mu\nu}_a}
-%   = \sum_{\v\rho} \sp{\prod_{\mu,\nu,a}\eta_{\mu\nu\rho^{\mu\nu}_a}}
-%   \bigotimes_{\mu,\nu,a} s_{\rho^{\mu\nu}_a},
-%   \label{eq:Q_K_rho}
-% \end{align}
-% where implicitly $\abs{\set{\rho^{\mu\nu}_a}}=r_{\mu\nu}$, as
-% $\p{\mu,\nu,a}$ essentially index a factor in \eqref{eq:Q_K_eta} and
-% $\rho^{\mu\nu}_a$ indexes one of the terms in that factor.  Letting
-% $\tilde\rho^{\mu\nu}_\kappa$ denote the number of elements in the
-% index list
-% $\p{\rho^{\mu\nu}_1,\rho^{\mu\nu}_2,\cdots,\rho^{\mu\nu}_{r_{\mu\nu}}}$
-% that are equal to $\kappa$, we observe that two terms in
-% \eqref{eq:Q_K_rho} with, say, $\v\rho=\v\rho_1$ and $\v\rho=\v\rho_2$
-% are equal up to a permutation of indices if
-% $\tilde{\v\rho}_1=\tilde{\v\rho}_2$.  The degeneracy (under
-% permutation) of terms in \eqref{eq:Q_K_rho}, i.e.~the number of
-% $\v\rho$ that are consistent with $\tilde{\v\rho}$, is
-% \begin{align}
-%   g_{\tilde{\v\rho}}
-%   \equiv \prod_{\mu,\nu,\kappa}
-%   { r_{\mu\nu} - \sum_{\lambda<\kappa} \tilde\rho^{\mu\nu}_\lambda
-%     \choose \tilde\rho^{\mu\nu}_\kappa }
-%   = \sp{\prod_{\mu,\nu} r_{\mu\nu}!}
-%   \sp{\prod_{\mu,\nu,\kappa} \tilde\rho^{\mu\nu}_\kappa!}^{-1}.
-% \end{align}
-% Defining additionally
-% \begin{align}
-%   \v\eta^{\tilde{\v\rho}}
-%   \equiv \prod_{\mu,\nu,\kappa}
-%   \p{\eta_{\mu\nu\kappa}}^{\tilde\rho^{\mu\nu}_\kappa},
-%   \label{eq:eta_rho}
-% \end{align}
-% we can collect like factors and group together equivalent terms in
-% \eqref{eq:Q_K_rho} to write
-% \begin{align}
-%   Q_{\v K}
-%   \simeq \sum_{\tilde{\v\rho}} g_{\tilde{\v\rho}} \v\eta^{\tilde{\v\rho}}
-%   \bigotimes_{\kappa}
-%   s_\kappa^{\otimes\sum_{\mu,\nu}\tilde\rho^{\mu\nu}_\kappa}
-%   \simeq \sum_{\tilde{\v\rho}} g_{\tilde{\v\rho}} \v\eta^{\tilde{\v\rho}}
-%   \1^{\otimes\sum_{\mu,\nu}\tilde\rho^{\mu\nu}_0}
-%   \otimes \bigotimes_{\kappa\ne0}
-%   s_\kappa^{\otimes\sum_{\mu,\nu}\tilde\rho^{\mu\nu}_\kappa},
-%   \label{eq:Q_K_rho_simp}
-% \end{align}
-% where again implicitly
-% $\sum_\kappa\tilde\rho^{\mu\nu}_\kappa=r_{\mu\nu}$ for consistency
-% with \eqref{eq:Q_K_eta} and \eqref{eq:Q_K_rho}, and we have explicitly
-% factored out the identity operators in each term of $Q_{\v K}$.
-
-% We have now simplified $Q_{\v K}$ sufficiently to substitute it back
-% into \eqref{eq:SS_PQ}, which (removing the tilde from $\tilde{\v\rho}$
-% to simplify notation) gives us
-% \begin{align}
-%   \tilde\S_{\v m} \tilde\S_{\v n}
-%   = \sum_{s\ge0} \sum_{\p{\v r;\v m,\v n,s}} f_{\v m\v n\v r}
-%   \sum_{\p{\v\rho;\v r}} g_{\v\rho} \v\eta^{\v\rho}
-%   c_{\v\ell_{\v m\v n\v r\v\rho}\v\rho}
-%   \tilde\S_{\v\ell_{\v m\v n\v r\v\rho}},
-%   \label{eq:SS_simp}
-% \end{align}
-% where the sum over $\p{\v\rho;\v r}$ denotes a sum over all values of
-% $\v\rho$ and with the restrictions
-% \begin{align}
-%   \abs{\set{\rho^{\mu\nu}_\kappa}} = \abs{\set{s_\kappa}},
-%   &&
-%   \sum_{\kappa} \rho^{\mu\nu}_\kappa = r_{\mu\nu};
-% \end{align}
-% the combinatorial factor
-% \begin{align}
-%   c_{\v\ell\v\rho}
-%   \equiv \f{\p{N - \abs{\v\ell}}!}
-%   {\p{N - \abs{\v\ell} - \sum_{\mu,\nu}\rho^{\mu\nu}_0}!}
-% \end{align}
-% accounts for the number of terms that are equivalent up to a
-% permutation of indices involving the identity operators in
-% \eqref{eq:Q_K_rho_simp}; and the number of $s_\mu$ operators in each
-% term of $\tilde\S_{\v\ell_{\v m\v n\v r\v\rho}}$ is
-% \begin{align}
-%   \ell_{\v m\v n\v r\v\rho;\mu}
-%   = m_\mu + n_\mu - \sum_\kappa \p{r_{\mu\kappa}+r_{\kappa\mu}}
-%   + \sum_{\a,\b} \rho^{\a\b}_\mu.
-% \end{align}
-
-% The sums in \eqref{eq:SS_simp} generally involve a large number of
-% terms; in practice, many of these terms will be equal to zero.  We
-% can, however, impose additional restrictions on these sums in such a
-% way as to throw out all terms that are zero, and keep only those that
-% are not.  Remembering that $s$ counts the number of spins that are
-% addressed by both operators $P_{\v j}$ and $P_{\v k}$ in
-% \eqref{eq:SS_PP}, the first restriction we can impose comes from
-% recognizing that for given $\tilde\S_{\v m},\tilde\S_{\v n}$, the
-% overlap $s$ is bounded as
-% \begin{align}
-%   \max\set{0,\abs{\v m}+\abs{\v n}-N}
-%   \le s \le \min\set{N,\abs{\v m}+\abs{\v n}}.
-% \end{align}
-% In addition, the matrix of structure constants $\eta_{\mu\nu\kappa}$
-% may generally contain zeros that we can preemptively avoid with
-% restrictions on $\v r$ and $\v\rho$.  Recalling that $r_{\mu\nu}$
-% counts the number of operators addressed by
-% $s_\mu s_\nu=\sum_\kappa\eta_{\mu\nu\kappa} s_\kappa$, if
-% $\eta_{\mu\nu\kappa}=0$ for all $\kappa$, i.e.~$s_\mu s_\nu=0$, then
-% any term with $r_{\mu\nu}>0$ will vanish.  Our second restriction on
-% the sums in \eqref{eq:SS_simp} is therefore to fix $r_{\mu\nu}=0$ for
-% all $\mu,\nu$ with $s_\mu s_\nu=0$.  Finally, we can avoid the
-% nullifying effect of zeros from the factor
-% $\v\eta^{\v\rho}=\prod_{\mu,\nu,\kappa}
-% \p{\eta_{\mu\nu\kappa}}^{\rho^{\mu\nu}_\kappa}$ by fixing
-% $\rho^{\mu\nu}_\kappa=0$ for all $\mu,\nu,\kappa$ with
-% $\eta_{\mu\nu\kappa}=0$.  These restrictions are sufficient to ensure
-% that all individual terms in \eqref{eq:SS_simp} are nonzero, although
-% they do not rule out the possibility of several terms canceling out.
-
-% TODO: comment on the use of this new basis vs.~the old one.
+\input{calculations_main.tex}
+
+\input{tutorial.tex}
+
+\input{calculations_OAT.tex}
+
+\input{noise_operators.tex}
+
+\input{linear_response.tex}
+
+% \input{changing_bases.tex}
 

### 2019-09-11 soc_squeezing/collective_paper/linear_response.tex
added comments on cumulant expansion added more discussion of the breakdown of TST added appendix on Heisenberg operators in open quantum systems, and the 'noise' operators

\section{Short-time linear response and two-time correlators}
\label{sec:linear_response}

Here we discuss the appearance of two-time correlation functions in
the short-time linear response of correlators to perturbations of a
Hamiltonian.  Consider an initial Hamiltonian $H$ perturbed by an
operator $V$ with $\norm{V}\ll\norm{H}$, where $\norm{\O}$ denotes the
operator norm of $\O$, such that the net Hamiltonian is
$\tilde H=H+V$.  We denote the generator of Heisenberg time evolution
under the perturbed (unperturbed) Hamiltonian by $\tilde T$ ($T$).
These generators are related by
\begin{align}
  \tilde T = T + i\ul V
\end{align}
where $\ul V$ is a superoperator whose action on operators $\O$ is
defined by
\begin{align}
  \ul V \O \equiv \sp{V,\O}_-.
\end{align}
Through quadratic order in the time $t$ and linear order in the
perturbation $\ul V$, we can say that
\begin{align}
  e^{t\tilde T}
  \approx \f12\sp{e^{tT}, e^{it\ul V}}_+
  \approx e^{tT} + \f12 it \sp{e^{t T}, \ul V}_+.
\end{align}
Defining perturbed and unperturbed Heisenberg operators
$\tilde\O\p{t}$ and $\O\p{t}$ that respectively satisfy
$\bk{\tilde\O\p{t}}=\bk{e^{t\tilde T}\O}$ and
$\bk{\O\p{t}}=\bk{e^{tT}\O}$, we thus find that for sufficiently small
times $t$ and weak perturbations $V$,
\begin{align}
  \bk{\tilde\O\p{t} - \O\p{t}}
  = \Bk{\p{e^{t\tilde T} - e^{tT}} \O}
  \approx \f12 i t \p{\bk{\sp{V,\O}_-\p{t}}
    + \bk{\sp{V,\O\p{t}}_-}}.
  \label{eq:response}
\end{align}
Two-time correlators $\bk{V\O\p{t}}$ and $\bk{\O\p{t}V}$, in addition
to the expectation values $\bk{\p{V\O}\p{t}}$ and
$\bk{\p{\O V}\p{t}}$, thus determine the short-time linear response of
correlators $\bk{\O\p{t}}$ to perturbations $V$ of a Hamiltonian.

### 2019-09-11 soc_squeezing/collective_paper/noise_operators.tex
added comments on cumulant expansion added more discussion of the breakdown of TST added appendix on Heisenberg operators in open quantum systems, and the 'noise' operators

\section{Heisenberg operators in open quantum systems}
\label{sec:noise}

Here we explain the origin and character of the mean-zero ``noise''
operators $\E_\O\p{t}$ that appear in the expansion of a Heisenberg
operator $\O\p{t}=\sum_{\Q\in\B}\O_\Q\p{t}\Q+\E_\O\p{t}$, where $\B$
denotes a complete basis of operators on the domain of
$\O\equiv\O\p{0}$ and $\O_\Q\p{t}$ are scalar coefficients.  Our
discussion should clarify why noise operators play no role in the
calculation of expectation values $\bk{\O\p{t}}$, as well as why noise
operators generally {\it do} need to be considered in the calculation
of OTOCs in open quantum systems, leading us to consider only the ``no
noise'' contribution to an OTOC for simplicity in Section
\ref{sec:multi_time}.

In any closed quantum system with initial state $\rho$ and propagator
$U\p{t}$, such that the state at time $t$ is
$\rho\p{t}\equiv U\p{t}\rho U\p{t}^\dag$, time-dependent Heisenberg
operators $\O\p{t}$ are uniquely defined from time-independent
Schr\"odinger operators $\O$ by
\begin{align}
  \bk{\O\p{t}} \equiv \tr\sp{\rho\p{t}\O} = \tr\sp{\rho\O\p{t}}.
  \label{eq:heisenberg_condition}
\end{align}
Enforcing \eqref{eq:heisenberg_condition} for {\it arbitrary} initial
states $\rho$ forces $\O\p{t}=U\p{t}^\dag\O U\p{t}$.  In an open
quantum system, however, the definition of a Heisenberg operator is
not so straightforward.  Open systems can often be understood as
subsystems of a larger closed system.  Consider therefore an open
system $S$ with environment $E$, a joint initial state $\rho_{SE}$,
and propagator $U_{SE}\p{t}$.  The reduced state $\rho_S\p{t}$ on $S$
at time $t$ is
\begin{align}
  \rho_S\p{t}
  \equiv \tr_E\sp{\rho_{SE}\p{t}}
  = \tr_E\sp{U_{SE}\p{t} \rho_{SE} U_{SE}\p{t}^\dag}
  \equiv \U_{\ul S}\p{t} \rho_S,
  \label{eq:state_S}
\end{align}
where $\rho_S\equiv\rho_S\p{0}$, $\ul S$ denotes the space of
operators on $S$, and the quantum channel $\U_{\ul S}\p{t}$ has the
decomposition\cite{rivas2012time}
\begin{align}
  \U_{\ul S}\p{t} \rho_S
  = \sum_j \U_S^{(j)}\p{t} \rho_S \U_S^{(j)}\p{t}^\dag
\end{align}
with ordinary operators $\U_S^{(j)}\p{t}$ on $S$.  We can thus expand
\begin{align}
  \bk{\O_S\p{t}}
  = \tr\sp{\rho_S\p{t}\O_S}
  = \tr\sp{\U_{\ul S}\p{t} \rho_S \O_S}
  = \tr\sp{\rho_S\p{t} \ul{\O_S}\p{t}},
\end{align}
where
\begin{align}
  \ul{\O_S}\p{t}
  \equiv \U_{\ul S}\p{t}^\dag \O_S
  = \sum_j \U_S^{(j)}\p{t}^\dag \O_S \U_S^{(j)}\p{t}.
  \label{eq:heisenberg_wrong}
\end{align}
We thus find that substituting $\O_S\p{t}\leftarrow\ul{\O_S}\p{t}$
suffices for the calculation of correlators $\bk{\O_S\p{t}}$, thereby
accounting for the validity of the equation of motion in
\eqref{eq:EOM}.

The problem with {\it defining} Heisenberg operators $\O_S\p{t}$ by
$\ul{\O_S}\p{t}$ only becomes evident when considering products of
Heisenberg operators.  One would like for the product of two
Heisenberg operators $\O_S\p{t}$ and $\Q_S\p{t}$ to satisfy
$\O_S\p{t}\Q_S\p{t}=\p{\O_S\Q_S}\p{t}$.  This intuition can be
formalized by observing that
\begin{align}
  \bk{\O_S\p{t}}
  = \tr\sp{\rho_{SE}\p{t} \p{\O_S\otimes\1_E}}
  = \tr\sp{\rho_{SE} \p{\O_S\otimes\1_E}\p{t}}
  = \bk{\p{\O_S\otimes\1_E}\p{t}},
\end{align}
where $\1_E$ is the identity operator on $E$, expectation values of
operators on system $A\in\set{S,E,SE}$ are taken with respect to
$\rho_A$, and
\begin{align}
  \p{\O_S\otimes\1_E}\p{t}
  \equiv U_{SE}\p{t}^\dag \p{\O_S\otimes\1_E} U_{SE}\p{t}.
  \label{eq:heisenberg_extension}
\end{align}
By expanding Heisenberg operators according to
\eqref{eq:heisenberg_extension}, we then find
\begin{align}
  \bk{\O_S\p{t}\Q_S\p{t}}
  = \bk{\p{\O_S\otimes\1_E}\p{t} \p{\Q_S\otimes\1_E}\p{t}}
  = \bk{\p{\O_S\Q_S\otimes\1_E}\p{t}}
  = \bk{\p{\O_S\Q_S}\p{t}}.
\end{align}
The expression in \eqref{eq:heisenberg_wrong}, however, makes it clear
that generally
$\ul{\O_S}\p{t} \ul{\Q_S}\p{t}\ne\p{\ul{\O_S\Q_S}}\p{t}$.  To correct
for this discrepancy, we define
\begin{align}
  \O_S\p{t} \equiv \ul{\O_S}\p{t} + \E_{\O_S}\p{t}
  \label{eq:noise_def}
\end{align}
in terms of new ``noise'' operators $\E_{\O_S}\p{t}$ that are
essentially defined to enforce the consistency of products such as
$\O_S\p{t}\Q_S\p{t}=\p{\O_S\Q_S}\p{t}$.  Self-consistency forces noise
operators to be mean-zero, as
\begin{align}
  \bk{\E_{\O_S}\p{t}} = \bk{\O_S\p{t}} - \bk{\ul{\O_S}\p{t}} = 0.
\end{align}
Furthermore, if we assume that the environment $E$ is Markovian, such
that to a good approximation
\begin{align}
  \rho_{SE}\p{t} = U_{SE}\p{t} \rho_{SE} U_{SE}\p{t}^\dag
  \approx \rho_S\p{t}\otimes\rho_E
  = \U_{\ul S}\p{t} \rho_S\otimes\rho_E,
  \label{eq:markov}
\end{align}
where $\rho_E$ is a steady state of the environment, then enforcing
\eqref{eq:markov} for all states $\rho_S$ implies
\begin{align}
  U_{SE}\p{t} \rho_{SE} \p{\O_S\otimes\1_E} U_{SE}\p{t}^\dag
  \approx \U_{\ul S}\p{t}\p{\rho_S\O_S}\otimes\rho_E.
  \label{eq:makov_op}
\end{align}
We can therefore expand
\begin{align}
  \bk{\O_S\Q_S\p{t}}
  &= \tr\sp{\rho_{SE} \p{\O_S\otimes\1_E}
    U_{SE}\p{t}^\dag \p{\Q_S\otimes\1_E} U_{SE}\p{t}} \\
  &= \tr\sp{U_{SE}\p{t} \rho_{SE} \p{\O_S\otimes\1_E}
    U_{SE}\p{t}^\dag \p{\Q_S\otimes\1_E}},
\end{align}
and invoke the Markov approximation in \eqref{eq:makov_op} to find
that
\begin{align}
  \bk{\O_S\Q_S\p{t}}
  \approx \tr\sp{\U_{\ul S}\p{t}\p{\rho_S\O_S} \Q_S}
  = \tr\sp{\rho_S \O_S \U_{\ul S}\p{t}^\dag \Q_S}
  = \bk{\O_S\ul{\Q_S}\p{t}}.
\end{align}
We thus find that for Markovian environments, noise operators play no
role in the calculation of correlators such as $C\p{t}$ in
\eqref{eq:two_time}, as
\begin{align}
  \bk{\O_S\E_{\Q_S}\p{t}}
  = \bk{\O_S\Q_S\p{t}} - \bk{\O_S\ul{\Q_S}\p{t}}
  \approx 0.
\end{align}
In contrast, noise operators generally {\it do} play a role in the
calculation of multi-time correlators of the form
$\bk{\prod_j\O_S^{(j)}\p{t_j}}$\cite{blocher2019quantum}.
Furthermore, these calculations generally require additional
assumptions about the environment.  To keep our discussion simple and
general, we therefore exclude the effects of noise terms in Section
\ref{sec:multi_time}.

### 2019-09-11 soc_squeezing/collective_paper/tutorial.tex
added comments on cumulant expansion added more discussion of the breakdown of TST added appendix on Heisenberg operators in open quantum systems, and the 'noise' operators

\section{Computing correlators with the truncated short-time (TST)
  expansion}
\label{sec:tutorial}

Here we provide a pedagogical tutorial for computing correlators using
the truncated short-time TST expansion.  For concreteness, we
nominally consider $N$ spins evolving under the one-axis twisting
(OAT) Hamiltonian
\begin{align}
  H_{\t{OAT}} = \chi S_\z^2,
\end{align}
additionally subject to spontaneous single-spin decay at rate
$\gamma_-$, with jump operators $\J_-=\set{s_-^{(j)}:j=1,2,\cdots,N}$.
The equation of motion for a Heisenberg operator
$S_+^\ell S_\z^m S_-^n$ is
\begin{align}
  \f{d}{dt} \bk{S_+^\ell S_\z^m S_-^n}
  = i\chi\Bk{\sp{S_\z^2, S_+^\ell S_\z^m S_-^n}_-}
  + \gamma_- \Bk{\D\p{\J_-} \p{S_+^\ell S_\z^m S_-^n}},
  \label{eq:tutorial_EOM}
\end{align}
where we can use the results in appendices \ref{sec:general_product}
and \ref{sec:decoherence_single} respectively to evaluate the
commutator $\sp{S_\z^2, S_+^\ell S_\z^m S_-^n}_-$ and dissipator
$\D\p{\J_-} \p{S_+^\ell S_\z^m S_-^n}$ in \eqref{eq:tutorial_EOM} (see
\eqref{eq:general_product} and \eqref{eq:decay_diff} in particular),
giving us
\begin{multline}
  \f{d}{dt} \bk{S_+^\ell S_\z^m S_-^n} \\
  = i\chi \bk{\p{\ell-n} S_+^\ell \p{\ell+n+2S_\z} S_\z^m S_-^n}
  + \gamma_- \Bk{S_+^\ell \sp{\p{S+S_\z}\p{-1+S_\z}^m
      - \p{S+\f{\ell+n}{2}+S_\z} S_\z^m} S_-^n}.
  \label{eq:tutorial_EOM_expanded}
\end{multline}
In practice, we do not want to keep track of such an expansion by
hand, especially in the case of e.g.~the two-axis twisting (TAT) and
twist-and-turn (TNT) models with more general types of decoherence,
for which the analogue of \eqref{eq:tutorial_EOM_expanded} may take
several lines just to write out in full.  Defining the operators
$\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$ with
$\v m\equiv\p{m_+,m_\z,m_-}$ for shorthand, we note that the vector
space spanned by $\set{\S_{\v m}}$ is closed under time evolution.  We
therefore expand
\begin{align}
  \f{d}{dt} \bk{\S_{\v n}}
  = \bk{T \S_{\v n}}
  = \sum_{\v m} \bk{\S_{\v m}} T_{\v m\v n},
  \label{eq:tutorial_EOM_general}
\end{align}
where $T$ is a superoperator that generates time evolution for
Heisenberg operators.  In the present example, the matrix elements
$T_{\v m\v n}\in\C$ of $T$ are defined by
\eqref{eq:tutorial_EOM_expanded} and \eqref{eq:tutorial_EOM_general}.
For any Hamiltonian $H$ with decoherence characterized by sets of jump
operators $\J$ and decoherence rates $\gamma_\J$, the matrix elements
$T_{\v m\v n}$ are more generally defined by
\begin{align}
  T \S_{\v n}
  = i\sp{H,\S_{\v n}}_-
  + \sum_\J \gamma_\J \D\p{\J} \S_{\v n}
  = \sum_{\v m} \S_{\v m} T_{\v m\v n}.
\end{align}
The results in Appendices \ref{sec:general_product},
\ref{sec:decoherence_single}, and \ref{sec:decoherence_collective} can
be used to write model-agnostic codes that compute matrix elements
$T_{\v m\v n}$, taking a particular Hamiltonian $H$ and decoherence
processes $\set{\p{\J,\gamma_\J}}$ as inputs.

In order to compute a quantity such as spin squeezing, we need to
compute correlators of the form $\bk{\S_{\v n}\p{t}}$, where
$\S_{\v n}\p{t}$ is the time-evolved Heisenberg operator corresponding
to $\S_{\v n}$, with $\S_{\v n}\p{0}\equiv\S_{\v n}$.  The order-$M$
truncated short-time (TST) expansion takes
\begin{align}
  \bk{\S_{\v n}\p{t}}
  = \sum_{k\ge0} \f{t^k}{k!} \bk{T^k\S_{\v n}}
  = \sum_{k\ge0} \f{t^k}{k!}
  \sum_{\v m} \bk{\S_{\v m}} T^k_{\v m\v n}
  \to \sum_{k=0}^M \f{t^k}{k!}
  \sum_{\v m} \bk{\S_{\v m}} T^k_{\v m\v n},
  \label{eq:tutorial_TST}
\end{align}
where $T^k_{\v m\v n}$ are matrix elements of the $k$-th time
derivative operator $T^k$, given by
\begin{align}
  T^0_{\v m\v n} \equiv
    \begin{cases}
      1 & \v m = \v n, \\
      0 & \t{otherwise}
    \end{cases},
  &&
  T^1_{\v m\v n} \equiv T_{\v m\v n},
  &&
  T^{k>1}_{\v m\v n}
  \equiv \sum_{\v p_1,\v p_2,\cdots,\v p_{k-1}}
  T_{\v m\v p_{k-1}} \cdots T_{\v p_3\v p_2}
  T_{\v p_2\v p_1} T_{\v p_1\v n}.
\end{align}
Matrix elements $T^k_{\v m\v n}$ and initial-time expectation values
$\bk{\S_{\v m}}$ are thus computed as needed for any particular
correlator $\bk{\S_{\v n}\p{t}}$ of interest, and combined according
to \eqref{eq:tutorial_TST}.  Note that initial-time expectation values
$\bk{\S_{\v m}}$ are an input to the TST expansion, and need to be
computed separately for any given initial state; expectation values
with respect to spin-polarized (Gaussian) states are provided in
Appendix \ref{sec:initial_conditions}.  In practice, we further
collect terms in \eqref{eq:tutorial_TST} to write
\begin{align}
  \bk{\S_{\v n}\p{t}} \to \sum_{k=0}^M c_{\v n k} t^k,
  &&
  c_{\v n k}
  \equiv \f1{k!} \sum_{\v m} \bk{\S_{\v m}} T^k_{\v m\v n},
  \label{eq:tutorial_corr}
\end{align}
where $c_{\v n k}$ are time-independent coefficients for the
polynomial time expansion of $\bk{\S_{\v n}\p{t}}$.  After computing
the coefficients $c_{\v n k}$, there is only negligible computational
overhead to compute the correlator $\bk{\S_{\v n}\p{t}}$ for any time
$t$ up to the breakdown time $t_M$ of an order-$M$ TST expansion.

### 2019-09-11 soc_squeezing/scripts/correlator_methods.py
simplified analytical expressions for OAT correlators

@@ -14,42 +14,42 @@ from special_functions import *
 # exact OAT results
 ##########################################################################################
 
-# exact correlators for OAT with decoherence; derivations in foss-feig2013nonequilibrium
+# exact correlators for OAT with decoherence
+# adapted from results in foss-feig2013nonequilibrium
 def correlators_OAT(spin_num, times, dec_rates):
     N = spin_num
-    SS = spin_num/2
+    S = N/2
     t = times
     D_p, D_z, D_m = dec_rates
 
-    gam = -(D_p - D_m) / 2
-    lam = (D_p + D_m) / 2
-    rr = D_p * D_m
-    Gam = D_z/2 + lam
+    gamma = ( D_p + D_m ) / 2
+    delta = D_p - D_m
+    kappa = sum(dec_rates)/2
 
     if D_m != 0 or D_p != 0:
         Sz_unit = (D_p-D_m)/(D_p+D_m) * (1-np.exp(-(D_p+D_m)*t))
     else:
         Sz_unit = np.zeros(len(times))
-    Sz = SS * Sz_unit
-    Sz_Sz = SS * (1/2 + (SS-1/2) * Sz_unit**2)
+    Sz = S * Sz_unit
+    Sz_Sz = S/2 + S*(S-1/2) * Sz_unit**2
 
-    def s(X): return X + 1j*gam
-    def sup_t_sinc(t,z): # e^(-\lambda t) t \sinc(t z)
-        if z != 0: return ( np.exp(-(lam-1j*z)*t) - np.exp(-(lam+1j*z)*t) ) / 2j / z
-        else: return np.exp(-lam*t) * t
-    def Phi(X):
-        z = np.sqrt(s(X)**2-rr)
-        val_cos = ( np.exp(-(lam-1j*z)*t) + np.exp(-(lam+1j*z)*t) ) / 2
-        val_sin = lam * sup_t_sinc(t,z)
-        return val_cos + val_sin
-    def Psi(X):
-        z = np.sqrt(s(X)**2-rr)
-        return (1j*s(X)-gam) * sup_t_sinc(t,z)
+    gamma_t = gamma*t
+    def sup_t_sinc(z): # e^{-gamma t} t \sinc(z t)
[…]

### 2019-09-11 soc_squeezing/scripts/correlator_methods.py
one more minor change to OAT analytics

@@ -34,17 +34,14 @@ def correlators_OAT(spin_num, times, dec_rates):
     Sz_Sz = S/2 + S*(S-1/2) * Sz_unit**2
 
     gamma_t = gamma*t
-    def sup_t_sinc(z): # e^{-gamma t} t \sinc(z t)
-        if z == 0: return np.exp(-gamma_t) * t
-        z_t = z * t
-        return ( np.exp(-gamma_t+1j*z_t) - np.exp(-gamma_t-1j*z_t) ) / 2j / z
-    def sup_cos(z): # e^{-\gamma t} \cos(z t)
-        z_t = z * t
-        return ( np.exp(-gamma_t+1j*z_t) + np.exp(-gamma_t-1j*z_t) ) / 2
+    def sup_cos(x): # e^{-\gamma t} \cos(x)
+        return ( np.exp(-gamma_t+1j*x) + np.exp(-gamma_t-1j*x) ) / 2
+    def sup_sin(x): # e^{-gamma t} \sin(x)
+        return ( np.exp(-gamma_t+1j*x) - np.exp(-gamma_t-1j*x) ) / 2j
 
     def w(x): return np.sqrt(x**2 - gamma**2 - 1j*x*delta)
-    def Phi(x): return sup_cos(w(x)) + gamma * sup_t_sinc(w(x))
-    def Psi(x): return (delta+1j*x) * sup_t_sinc(w(x))
+    def Phi(x): return sup_cos(w(x)*t) + gamma/w(x) * sup_sin(w(x)*t)
+    def Psi(x): return (delta+1j*x)/w(x) * sup_sin(w(x)*t)
 
     Sp = S * np.exp(-kappa*t) * Phi(1)**(N-1)
     Sp_Sz = -1/2 * Sp + S * (S-1/2) * np.exp(-kappa*t) * Psi(1) * Phi(1)**(N-2)

### 2019-09-12 soc_squeezing/collective_paper/collective_spin.tex
some changes to notation

@@
 where $\uv S\equiv\p{\hat S_\x,\hat S_\y,\hat S_\z}$ is a collective
-spin operator-valued vector, and the minimization is performed over
-all unit vectors $\v v$ orthogonal to the mean spin vector
-$\bk{\uv S}$.  This squeezing parameter is entirely determined by one-
-and two-spin correlators of the form $\bk{\hat S_\a}$ and
-$\bk{\hat S_\a \hat S_\b}$.  For the unitary dynamics discussed in
-this work, these correlators are obtainable via exact simulations of
-quantum dynamics in the $\p{N+1}$-dimensional Dicke manifold of states
-$\set{\ket{S,m}}$ with net spin $S$ and spin projection $m$ onto the
-$z$ axis, i.e.~with $\bk{S,m|\uv S^2|S,m}=S\p{S+1}$ and
-$\bk{S,m|\hat S_\z|S,m}=m$ for $m\in\set{-S,-S+1,\cdots,S}$.  In the
-presence of single-spin or collective decoherence, meanwhile, these
-correlators are obtainable with the collective-spin quantum trajectory
-Monte Carlo method developed in ref.~\cite{zhang2018montecarlo}.  In
-this work, these exact and quantum trajectory simulations will be used
-to benchmark the TST expansion in Eq.~\eqref{eq:TST}.
+spin operator-valued vector, the minimization is performed over all
+unit vectors $\v v$ orthogonal to the mean spin vector $\bk{\uv S}$,
+and for brevity we have suppressed the explicit time dependence of
+operators in Eq.~\eqref{eq:squeezing}.  This squeezing parameter is
+entirely determined by one- and two-spin correlators of the form
+$\bk{\hat S_\a}$ and $\bk{\hat S_\a \hat S_\b}$.  For the unitary
+dynamics discussed in this work, these correlators are obtainable via
+exact simulations of quantum dynamics in the $\p{N+1}$-dimensional
+Dicke manifold of states $\set{\ket{S,m}}$ with net spin $S$ and spin
+projection $m$ onto the $z$ axis, i.e.~with
+$\bk{S,m|\uv S^2|S,m}=S\p{S+1}$ and $\bk{S,m|\hat S_\z|S,m}=m$ for
+$m\in\set{-S,-S+1,\cdots,S}$.  In the presence of single-spin or
+collective decoherence, meanwhile, these correlators are obtainable
+with the collective-spin quantum trajectory Monte Carlo method
+developed in ref.~\cite{zhang2018montecarlo}.  In this work, these
+exact and quantum trajectory simulations will be used to benchmark the
+TST expansion in Eq.~\eqref{eq:TST}.
 

### 2019-09-12 soc_squeezing/collective_paper/noise_operators.tex
some changes to notation

@@
 Enforcing \eqref{eq:heisenberg_condition} for {\it arbitrary} initial
-states $\rho$ forces $\O\p{t}=U\p{t}^\dag\O U\p{t}$.  In an open
+states $\rho$ forces $\O\p{t}=U^\dag\p{t}\O U\p{t}$.  In an open
 quantum system, however, the definition of a Heisenberg operator is
@@
 system $S$ with environment $E$, a joint initial state $\rho_{SE}$,
-and propagator $U_{SE}\p{t}$.  The reduced state $\rho_S\p{t}$ on $S$
+and propagator $U_{SE}\p{t}$.  The reduced state $\rho_S\p{t}$ of $S$
 at time $t$ is
@@
   \equiv \tr_E\sp{\rho_{SE}\p{t}}
-  = \tr_E\sp{U_{SE}\p{t} \rho_{SE} U_{SE}\p{t}^\dag}
+  = \tr_E\sp{U_{SE}\p{t} \rho_{SE} U_{SE}^\dag\p{t}}
   \equiv \U_{\ul S}\p{t} \rho_S,
@@
   \U_{\ul S}\p{t} \rho_S
-  = \sum_j \U_S^{(j)}\p{t} \rho_S \U_S^{(j)}\p{t}^\dag
+  = \sum_j \U_S^{(j)}\p{t} \rho_S {\U_S^{(j)}}^\dag\p{t}
 \end{align}
@@
   \ul{\O_S}\p{t}
-  \equiv \U_{\ul S}\p{t}^\dag \O_S
-  = \sum_j \U_S^{(j)}\p{t}^\dag \O_S \U_S^{(j)}\p{t}.
+  \equiv \U_{\ul S}^\dag\p{t} \O_S
+  = \sum_j {\U_S^{(j)}}^\dag\p{t} \O_S \U_S^{(j)}\p{t}.
   \label{eq:heisenberg_wrong}
 \end{align}
-We thus find that substituting $\O_S\p{t}\leftarrow\ul{\O_S}\p{t}$
-suffices for the calculation of correlators $\bk{\O_S\p{t}}$, thereby
-accounting for the validity of the equation of motion in
-\eqref{eq:EOM}.
+We thus find that substituting $\ul{\O_S}\p{t}$ in place of
+$\O_S\p{t}$ suffices for the calculation of correlators
+$\bk{\O_S\p{t}}$, thereby accounting for the validity of the equation
+of motion in \eqref{eq:EOM}.
 
@@
 where $\1_E$ is the identity operator on $E$, expectation values of
-operators on system $A\in\set{S,E,SE}$ are taken with respect to
-$\rho_A$, and
+Heisenberg operators on system $A\in\set{S,E,SE}$ are taken with
+respect to the state $\rho_A$, and
 \begin{align}
   \p{\O_S\otimes\1_E}\p{t}
-  \equiv U_{SE}\p{t}^\dag \p{\O_S\otimes\1_E} U_{SE}\p{t}.
+  \equiv U_{SE}^\dag\p{t} \p{\O_S\otimes\1_E} U_{SE}\p{t}.
   \label{eq:heisenberg_extension}
@@
 \begin{align}
-  \rho_{SE}\p{t} = U_{SE}\p{t} \rho_{SE} U_{SE}\p{t}^\dag
+  \rho_{SE}\p{t} = U_{SE}\p{t} \rho_{SE} U_{SE}^\dag\p{t}
   \approx \rho_S\p{t}\otimes\rho_E
@@
 \end{align}
-where $\rho_E$ is a steady state of the environment, then enforcing
-\eqref{eq:markov} for all states $\rho_S$ implies
+with $\rho_E$ a time-independent steady state of the environment, then
+enforcing \eqref{eq:markov} for all states $\rho_S$ implies
 \begin{align}
-  U_{SE}\p{t} \rho_{SE} \p{\O_S\otimes\1_E} U_{SE}\p{t}^\dag
+  U_{SE}\p{t} \rho_{SE} \p{\O_S\otimes\1_E} U_{SE}^\dag\p{t}
   \approx \U_{\ul S}\p{t}\p{\rho_S\O_S}\otimes\rho_E.
@@
   &= \tr\sp{\rho_{SE} \p{\O_S\otimes\1_E}
-    U_{SE}\p{t}^\dag \p{\Q_S\otimes\1_E} U_{SE}\p{t}} \\
+    U_{SE}^\dag\p{t} \p{\Q_S\otimes\1_E} U_{SE}\p{t}} \\
   &= \tr\sp{U_{SE}\p{t} \rho_{SE} \p{\O_S\otimes\1_E}
-    U_{SE}\p{t}^\dag \p{\Q_S\otimes\1_E}},
+    U_{SE}^\dag\p{t} \p{\Q_S\otimes\1_E}},
 \end{align}
@@
   \approx \tr\sp{\U_{\ul S}\p{t}\p{\rho_S\O_S} \Q_S}
-  = \tr\sp{\rho_S \O_S \U_{\ul S}\p{t}^\dag \Q_S}
+  = \tr\sp{\rho_S \O_S \U_{\ul S}^\dag\p{t} \Q_S}
   = \bk{\O_S\ul{\Q_S}\p{t}}.

### 2019-09-12 soc_squeezing/collective_paper/tutorial.tex
some changes to notation

@@
 The equation of motion for a Heisenberg operator
-$S_+^\ell S_\z^m S_-^n$ is
+$\p{S_+^\ell S_\z^m S_-^n}\p{t}$ is
 \begin{align}
-  \f{d}{dt} \bk{S_+^\ell S_\z^m S_-^n}
+  \f{d}{dt} \Bk{S_+^\ell S_\z^m S_-^n}
   = i\chi\Bk{\sp{S_\z^2, S_+^\ell S_\z^m S_-^n}_-}
-  + \gamma_- \Bk{\D\p{\J_-} \p{S_+^\ell S_\z^m S_-^n}},
+  + \gamma_- \Bk{\D\p{\J_-}\p{S_+^\ell S_\z^m S_-^n}},
   \label{eq:tutorial_EOM}
 \end{align}
-where we can use the results in appendices \ref{sec:general_product}
-and \ref{sec:decoherence_single} respectively to evaluate the
-commutator $\sp{S_\z^2, S_+^\ell S_\z^m S_-^n}_-$ and dissipator
-$\D\p{\J_-} \p{S_+^\ell S_\z^m S_-^n}$ in \eqref{eq:tutorial_EOM} (see
-\eqref{eq:general_product} and \eqref{eq:decay_diff} in particular),
-giving us
+where we have suppressed the explicit time dependence of operators for
+brevity.  Using the results in appendices \ref{sec:general_product}
+and \ref{sec:decay_single} respectively to evaluate the commutator
+$\sp{S_\z^2, S_+^\ell S_\z^m S_-^n}_-$ and dissipator
+$\D\p{\J_-} \p{S_+^\ell S_\z^m S_-^n}$ in \eqref{eq:tutorial_EOM}, we
+can expand
 \begin{multline}
@@
 In order to compute a quantity such as spin squeezing, we need to
-compute correlators of the form $\bk{\S_{\v n}\p{t}}$, where
-$\S_{\v n}\p{t}$ is the time-evolved Heisenberg operator corresponding
-to $\S_{\v n}$, with $\S_{\v n}\p{0}\equiv\S_{\v n}$.  The order-$M$
-truncated short-time (TST) expansion takes
+compute correlators of the form $\bk{\S_{\v n}\p{t}}$, where for
+clarity we will re-introduce the explicit time dependence of
+Heisenberg operators $\S_{\v n}\p{t}$.  The order-$M$ truncated
+short-time (TST) expansion takes
 \begin{align}
   \bk{\S_{\v n}\p{t}}
-  = \sum_{k\ge0} \f{t^k}{k!} \bk{T^k\S_{\v n}}
+  = \bk{e^{tT}\S_{\v n}\p{0}}
+  = \sum_{k\ge0} \f{t^k}{k!} \bk{T^k\S_{\v n}\p{0}}
   = \sum_{k\ge0} \f{t^k}{k!}
-  \sum_{\v m} \bk{\S_{\v m}} T^k_{\v m\v n}
+  \sum_{\v m} \bk{\S_{\v m}\p{0}} T^k_{\v m\v n}
   \to \sum_{k=0}^M \f{t^k}{k!}
-  \sum_{\v m} \bk{\S_{\v m}} T^k_{\v m\v n},
+  \sum_{\v m} \bk{\S_{\v m}\p{0}} T^k_{\v m\v n},
   \label{eq:tutorial_TST}
@@
 Matrix elements $T^k_{\v m\v n}$ and initial-time expectation values
-$\bk{\S_{\v m}}$ are thus computed as needed for any particular
+$\bk{\S_{\v m}\p{0}}$ are thus computed as needed for any particular
 correlator $\bk{\S_{\v n}\p{t}}$ of interest, and combined according
 to \eqref{eq:tutorial_TST}.  Note that initial-time expectation values
-$\bk{\S_{\v m}}$ are an input to the TST expansion, and need to be
-computed separately for any given initial state; expectation values
-with respect to spin-polarized (Gaussian) states are provided in
-Appendix \ref{sec:initial_conditions}.  In practice, we further
-collect terms in \eqref{eq:tutorial_TST} to write
+$\bk{\S_{\v m}\p{0}}$ are an {\it input} to the TST expansion, and
+need to be computed separately for any initial state of interest;
+expectation values with respect to spin-polarized (Gaussian) states
+are provided in Appendix \ref{sec:initial_conditions}.  In practice,
+we further collect terms in \eqref{eq:tutorial_TST} to write
 \begin{align}
@@
   c_{\v n k}
-  \equiv \f1{k!} \sum_{\v m} \bk{\S_{\v m}} T^k_{\v m\v n},
+  \equiv \f1{k!} \sum_{\v m} \bk{\S_{\v m}\p{0}} T^k_{\v m\v n},
   \label{eq:tutorial_corr}
 \end{align}
-where $c_{\v n k}$ are time-independent coefficients for the
-polynomial time expansion of $\bk{\S_{\v n}\p{t}}$.  After computing
-the coefficients $c_{\v n k}$, there is only negligible computational
-overhead to compute the correlator $\bk{\S_{\v n}\p{t}}$ for any time
-$t$ up to the breakdown time $t_M$ of an order-$M$ TST expansion.
+where $c_{\v n k}$ are time-independent coefficients for the expansion
+of $\bk{\S_{\v n}\p{t}}$.  After computing the coefficients
+$c_{\v n k}$, there is only negligible computational overhead to
+compute the correlator $\bk{\S_{\v n}\p{t}}$ for any time $t$.

### 2019-09-12 soc_squeezing/collective_paper/collective_spin.tex
minor additions/additions to text concerning Heisenberg operators

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{document}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Theory}
@@
 self-adjoint, for example, then $\O_{\v m}^*=\O_{\v m^*}$ with
-$\v m^*\equiv\p{m_-,m_\z,m_+}$.
-%
-% The corresponding Heisenberg operator is then
-% $\hat\O\p{t}=\sum_{\v m}\O_{\v m}\p{t}\hat\S_{\v m}$, with
-% time-dependent coefficients $\O_{\v n}\p{t}$ for time-independent
-% (Schr\"odinger) operators $\hat\S_{\v m}$.  The time evolution of
-% Heisenberg operators is determined by
-%
-The corresponding Heisenberg operator is then
+$\v m^*\equiv\p{m_-,m_\z,m_+}$.  The corresponding Heisenberg operator
+is then
 $\hat\O\p{t}=\sum_{\v m}\O_{\v m}\p{t}\hat\S_{\v m}+\hat\E_\O\p{t}$,
-with time-dependent coefficients $\O_{\v n}\p{t}$ for time-independent
-(Schr\"odinger) operators $\hat\S_{\v m}$, and mean-zero ``noise''
+with time-dependent coefficients $\O_{\v m}\p{t}$ for time-independent
+Schr\"odinger operators $\hat\S_{\v m}$, and mean-zero ``noise''
 operators $\hat\E_\O\p{t}$ that result from interactions between the
@@
 evolve according to
-%
 \begin{align}
@@
 \end{align}
-%
-% We refer Eq.~\eqref{eq:TST} as the truncated short-time (TST)
-% expansion of Heisenberg operators.  Unlike the truncation in
-% Eq.~\eqref{eq:weight_truncation}, the nonzero matrix elements
-% $T^k_{\v m\v n}$ for $k=0,1,\cdots,M$ in Eq.~\eqref{eq:TST} tell us
-% which operators are relevant for computing the expectation value
-% $\bk{\hat\O\p{t}}$ to a fixed order $M$.  Note that using the
-% relation $\hat\S_{\v m}^\dag=\hat\S_{\v m^*}$, which by Hermitian
-% conjugation of Eq.~\eqref{eq:EOM} also implies that
-% $T_{\v m^*\v n^*}=T_{\v m\v n}^*$, can cut both the number of
-% initial-time expectation values $\bk{\hat\S_{\v m}}_{t=0}$ and the
-% number of matrix elements $T_{\v m\v n}$ that we may need to
-% explicitly compute roughly in half.
-%
 We refer Eq.~\eqref{eq:TST} as the truncated short-time (TST)
@@
 correlators using the TST expansion in Appendix \ref{sec:tutorial}.
-%
 
-%
-% In principle, the factorial suppression of terms at higher orders of
-% the expansion in Eq.~\eqref{eq:time_series} implies for any given
-% time $t$, there exists a truncation order $M_t$ for which the
-% truncation error in Eq.~\eqref{eq:TST} is negligibly small.  In
-% practice, only a maximal truncation order $M_{\t{max}}$ is
-% accessible with limited computational resources, such that the TST
-% expansion in Eq.~\eqref{eq:TST} only allows us to compute the
-% expectation value $\bk{\hat\O\p{t}}$ to a maximal time
-% $t_{M_{\t{max}}}$.  As we will see in the following section,
-% $M_{\t{max}}=35$ will suffice for the computation of collective spin
-% correlators up to times that are relevant e.g.~for spin squeezing
-% protocols.  Appendix \ref{sec:tutorial} provides a pedagogical
-% tutorial for computing correlators using the TST expansion.
-%
 Three primary considerations limit the maximum time $t$ to which we
@@
 in this work.
-%
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Spin squeezing, benchmarking, and breakdown}
@@
 $\ket\X\to\ket{-\Z}\equiv\ket\dn^{\otimes N}$, as well as appropriate
-transformations of the Hamiltonian and jump operators.
-%
-% The only effect of this rotation on the results presented in Figure
-% \ref{fig:benchmarking} is to prolong the time for which the TST
-% expansion agrees with the benchmarking simulations.  The reason for
-% different results in rotated basis has to do with the breakdown of
-% the TST expansion, which we discuss further below.
-%
-The only effects of this rotation on the results presented in Figure
+transformations of the Hamiltonian and jump operators.  The only
+effects of this rotation on the results presented in Figure
 \ref{fig:benchmarking} are to
@@
 breakdown to future work.
-%
 
@@
 individual correlators taking unphysical values with
-$\abs{\bk{\hat\S_{\v n}\p{t}}}\gtrsim S^{\abs{\v n}}$.
-%
-% The sudden and drastic departure from virtually exact results occurs
-% because the TST expansion neglects high-weight
-% (i.e.~large-$\abs{\v m}$) operators $\hat\S_{\v m}$ whose
-% contributions to a Heisenberg operator $\hat\O\p{t}$ of interest
-% eventually become non-negligible.  For sufficiently large times $t$,
-% these individual contributions can become comparable in magnitude to
-% the expectation value $\bk{\hat\O\p{t}}$, and will grow with some
-% large power of $t$, so their truncation will have a drastic effect
-% on the estimated value of $\bk{\hat\O\p{t}}$.  This breakdown
-% mechanism is also the reason for prolonged agreement between the TST
-% expansion and benchmarking methods when squeezing the initial state
-% $\ket{-\Z}$ rather than $\ket{\X}$: for initial states $\ket{-\Z}$,
-% all initial-time correlators $\bk{\hat\S_{\v m}}$ vanish unless
-% $m_+=m_-=0$ (see Appendix \ref{sec:initial_conditions}), so there is
-% a substantially smaller number of neglected terms with non-zero
-% contribution to the correlator $\bk{\hat\O\p{t}}$.
-%
-Going up through order $M=70$ in the TST expansion does not
-significantly increase the breakdown time $t_{\t{break}}^{(M)}$, and
-in some cases even shortens $t_{\t{break}}^{(M)}$.  The sudden and
-drastic departure from virtually exact results is consistent with the
-limitations of the TST expansion discussed at the end of Section
-\ref{sec:theory}.  Specifically, we identify three possible mechanisms
-for breakdown:
+$\abs{\bk{\hat\S_{\v n}\p{t}}}\gtrsim S^{\abs{\v n}}$.  Going up
+through order $M=70$ in the TST expansion does not significantly
+increase the breakdown time $t_{\t{break}}^{(M)}$, and in some cases
+even shortens $t_{\t{break}}^{(M)}$.  The sudden and drastic departure
+from virtually exact results is consistent with the limitations of the
+TST expansion discussed at the end of Section \ref{sec:theory}.
+Specifically, we identify three possible mechanisms for breakdown:
 \begin{enumerate*}
@@
 correlators.
-%
 
-%
-% Although the TST expansion breaks down at short times, it has two
-% key advantages over other methods to compute collective spin
-% correlators.  First, computing spin correlators with the TST
-% expansion is generally much faster and requires much fewer computing
-% resources than the alternatives.  The quantum trajectory Monte Carlo
-% simulations performed for Figure \ref{fig:squeezing_incoherent}, for
-% example, take $\sim10^4$ CPU hours to compute on standard modern
-% computing hardware; the bulk of this time is spent performing sparse
-% matrix-vector multiplication, leaving little room to further
-% optimize serial runtime.  Parallelization can reduce the actual
-% runtime of these quantum trajectory simulations to $\sim10$ hours by
-% running all trajectories at once, but at the cost of greatly
-% increasing computing resource requirements.  The TST expansion
-% results in Figure \ref{fig:squeezing_incoherent}, meanwhile, take
-% $\sim10$ seconds to compute with a single CPU on similar hardware.
-%
 Although the TST expansion breaks down at short times, it has two key
@@
 \end{enumerate*}
-%
 
@@
 dephasing of individual spins at rates
-$\gamma_-=\gamma_+=\gamma_\z=100\chi$.
-%
-% These results show that
-%
-The system size in these simulations is too large for straightforward
-application of exact methods for open quantum systems.  Quantum
-trajectory Monte Carlo simulations, meanwhile, take a prohibitively
-long time to converge with such strong decoherence due to the
-multiplicity of quantum trajectories that require averaging.
+$\gamma_-=\gamma_+=\gamma_\z=100\chi$.  The system size in these
+simulations is too large for straightforward application of exact
+methods for open quantum systems.  Quantum trajectory Monte Carlo
+simulations, meanwhile, take a prohibitively long time to converge
+with such strong decoherence due to the multiplicity of quantum
+trajectories that require averaging.
 
-The results in Figure \ref{fig:decoherence_strong} show that
-%
-TNT is much more robust to single-spin decoherence than OAT or TAT, a
-finding that could not necessarily be deduced from the
-weak-decoherence simulations presented in Figure
-\ref{fig:benchmarking}.  Strong-decoherence computations of this sort
-were used to put lower bounds on theoretically achievable spin
-squeezing via TAT with decoherence in
-Ref.~[\citenum{he2019engineering}], exemplifying a concrete and
-practical application of the TST expansion and the collective-spin
-structure constants calculated in this work.
+The results in Figure \ref{fig:decoherence_strong} show that TNT is
+much more robust to single-spin decoherence than OAT or TAT, a finding
+that could not necessarily be deduced from the weak-decoherence
+simulations presented in Figure \ref{fig:benchmarking}.
+Strong-decoherence computations of this sort were used to put lower
+bounds on theoretically achievable spin squeezing via TAT with
+decoherence in Ref.~[\citenum{he2019engineering}], exemplifying a
+concrete and practical application of the TST expansion and the
+collective-spin structure constants calculated in this work.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Two-time correlation functions and out-of-time-ordered
@@
 in the context of the squeezing models in Section \ref{sec:squeezing}.
-%
 The subscript on $\bk{\cdot}_{\t{nn}}$ in Eq.~\eqref{eq:four_point}
@@
 underlying the calculation of multi-time correlators.
-%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Conclusions}

### 2019-09-12 soc_squeezing/collective_paper/noise_operators.tex
minor additions/additions to text concerning Heisenberg operators

@@
 operators $\E_\O\p{t}$ that appear in the expansion of a Heisenberg
-operator $\O\p{t}=\sum_{\Q\in\B}\O_\Q\p{t}\Q+\E_\O\p{t}$, where $\B$
-denotes a complete basis of operators on the domain of
-$\O\equiv\O\p{0}$ and $\O_\Q\p{t}$ are scalar coefficients.  Our
-discussion should clarify why noise operators play no role in the
-calculation of expectation values $\bk{\O\p{t}}$, as well as why noise
-operators generally {\it do} need to be considered in the calculation
-of OTOCs in open quantum systems, leading us to consider only the ``no
-noise'' contribution to an OTOC for simplicity in Section
-\ref{sec:multi_time}.
+operator $\O\p{t}=\sum_{\v m}\O_{\v m}\p{t}\S_{\v m}+\E_\O\p{t}$ with
+time-dependent coefficients $\O_{\v m}\p{t}$ for time-independent
+(Schr\"odinger) operators $\hat\S_{\v m}$.  Our discussion clarify why
+noise operators play no role in our calculation of expectation values
+of the form $\bk{\O\p{t}}$ and $\bk{\O\Q\p{t}}$, despite the fact that
+noise operators generally {\it do} need to be considered in the
+calculation of more general multi-time correlators in open quantum
+systems\cite{blocher2019quantum}.
 
@@
 $U\p{t}$, such that the state at time $t$ is
-$\rho\p{t}\equiv U\p{t}\rho U\p{t}^\dag$, time-dependent Heisenberg
+$\rho\p{t}\equiv U\p{t}\rho U^\dag\p{t}$, time-dependent Heisenberg
 operators $\O\p{t}$ are uniquely defined from time-independent
@@
 \end{align}
-where $\rho_S\equiv\rho_S\p{0}$, $\ul S$ denotes the space of
-operators on $S$, and the quantum channel $\U_{\ul S}\p{t}$ has the
+where $\rho_S\equiv\rho_S\p{0}$ is a time-independent state of $S$ in
+the Heisenberg picture, $\ul S$ denotes the space of operators on $S$,
+and the quantum channel $\U_{\ul S}\p{t}$ has the
 decomposition\cite{rivas2012time}
@@
 \end{align}
-with ordinary operators $\U_S^{(j)}\p{t}$ on $S$.  We can thus expand
+with ordinary operators $\U_S^{(j)}\p{t}$ on $S$.  We can therefore
+expand
 \begin{align}
@@
   = \tr\sp{\U_{\ul S}\p{t} \rho_S \O_S}
-  = \tr\sp{\rho_S\p{t} \ul{\O_S}\p{t}},
+  = \tr\sp{\rho_S \U_{\ul S}^\dag\p{t} \O_S}
+  = \tr\sp{\rho_S \ul{\O_S}\p{t}}
+  = \bk{\ul{\O_S}\p{t}},
+  \label{eq:heisenberg_correlator}
 \end{align}
-where
+where we define the operator
 \begin{align}
@@
 $\bk{\O_S\p{t}}$, thereby accounting for the validity of the equation
-of motion in \eqref{eq:EOM}.
+of motion in \eqref{eq:EOM}.  As we show below, this substitution also
+suffices for the calculation of two-time correlators of the form
+$\bk{\O_S\Q_S\p{t}}$ when the environment $E$ is Markovian.
 
@@
   = \bk{\p{\O_S\otimes\1_E}\p{t}},
+  \label{eq:heisenberg_expansion}
 \end{align}
@@
 \end{align}
-By expanding Heisenberg operators according to
-\eqref{eq:heisenberg_extension}, we then find
+By expanding Heisenberg operators similarly to
+\eqref{eq:heisenberg_expansion} and \eqref{eq:heisenberg_extension},
+we then find
 \begin{align}
@@
 in terms of new ``noise'' operators $\E_{\O_S}\p{t}$ that are
-essentially defined to enforce the consistency of products such as
-$\O_S\p{t}\Q_S\p{t}=\p{\O_S\Q_S}\p{t}$.  Self-consistency forces noise
-operators to be mean-zero, as
+essentially defined to enforce the consistency of operator products
+such as $\O_S\p{t}\Q_S\p{t}=\p{\O_S\Q_S}\p{t}$. Self-consistency
+forces noise operators to be mean-zero, as
 \begin{align}
@@
 \end{align}
-Furthermore, if we assume that the environment $E$ is Markovian, such
-that to a good approximation
+Furthermore, if the environment $E$ is Markovian, then noise operators
+are also uncorrelated with initial-time observables,
+i.e.~$\bk{\O_S\E_{\Q_S}\p{t}}=0$, which means that noise operators can
+be neglected in the calculation of two-time correlators of the form
+$\bk{\O_S\Q_S\p{t}}$.  To see why, we observe that a Markovian
+environment is essentially defined to satisfy
 \begin{align}
@@
 \end{align}
-with $\rho_E$ a time-independent steady state of the environment, then
-enforcing \eqref{eq:markov} for all states $\rho_S$ implies
+with $\rho_E$ a time-independent steady state of the environment.
+Enforcing \eqref{eq:markov} for all states $\rho_S$ implies that
 \begin{align}
@@
   = \tr\sp{\rho_S \O_S \U_{\ul S}^\dag\p{t} \Q_S}
-  = \bk{\O_S\ul{\Q_S}\p{t}}.
+  = \bk{\O_S\ul{\Q_S}\p{t}},
 \end{align}
-We thus find that for Markovian environments, noise operators play no
-role in the calculation of correlators such as $C\p{t}$ in
-\eqref{eq:two_time}, as
+which implies
 \begin{align}
@@
 \end{align}
-In contrast, noise operators generally {\it do} play a role in the
-calculation of multi-time correlators of the form
+Noise operators thus play no role in the calculation of correlators
+such as $C\p{t}$ in \eqref{eq:two_time}.  In contrast, noise operators
+generally {\it do} play a role in the calculation of multi-time
+correlators of the form
 $\bk{\prod_j\O_S^{(j)}\p{t_j}}$\cite{blocher2019quantum}.

### 2019-09-13 soc_squeezing/squeezing_paper/main.tex
uploading final changes prior to last submission of squeezing paper

@@
 a state that is squeezed with respect to spin operators that are homogeneous (i.e.~of the form $S_\x,S_\y,S_\z$) in a particular gauge is not necessarily squeezed with respect to spin operators that are homogeneous in a different gauge.
-In this appendix, we will work explicitly in the ``lab gauge'' of the clock interrogation protocol, in which the Fermi-Hubbard Hamiltonian is SOC-free and the $\phi=0$ clock laser is homogeneous.
+In this analysis, we will work explicitly in the ``lab gauge'' of the clock interrogation protocol, in which the Fermi-Hubbard Hamiltonian is SOC-free and the $\phi=0$ clock laser is homogeneous.
 To resolve the fact that our squeezing protocol prepares a state that is squeezed in the ``wrong gauge'', we will construct a simple pulse sequence that transforms the inhomogeneous spin operators accessible by the $\phi\ne0$ laser into a homogeneous form in our lab gauge.

### 2019-09-13 soc_squeezing/squeezing_paper/supplementary.tex
uploading final changes prior to last submission of squeezing paper

@@
 \usepackage{graphicx} % for figures
-\graphicspath{{./figures/}} % set path for all figures
+\graphicspath{{../figures/}} % set path for all figures
 
@@
 Current 3D optical lattice implementations involve a harmonic confining potential, which significantly alters the underlying single-particle eigenstructure and can potentially degrade accessible squeezing within our protocol.
-In this appendix, we examine the effect of a harmonic trap on our protocol and discuss strategies to mitigate undesired effects.
+In this section of the supplement, we examine the effect of a harmonic trap on our protocol and discuss strategies to mitigate undesired effects.
 We model the trap by the addition of the term
@@
 The big advantage here compared to  prior experiments done in a 1D lattice at $\mu$K temperature\cite{martin2013quantum} is that in a Fermi  degenerate  gas loaded in a 3D optical lattice, $p$-wave losses are further suppressed by the centrifugal barrier and Pauli blocking, and only happen through a wave-function overlap between atoms at different lattice sites.
-In this appendix, we quantify the effect of $p$-wave interactions on squeezing.
+In this supplementary material, we quantify the effect of $p$-wave interactions on squeezing.
 To account for $p$-wave losses, we describe the dynamics  using a master equation for the system's density matrix $\hat{\rho}$:

### 2019-09-13 multi_body_interactions/papers/theory/paper.tex
fixed minor typo in multi-body theory paper

@@
 ref.~\cite{zhang2014spectroscopic}.  The squared relative momentum
-operator $k_z^2$ is represented by symmetrized left- and right-acting
-derivative operators in order to preserve manifest Hermiticity of
-$H_{\t{int}}'$.  At third order in the low-energy effective theory
-developed in section \ref{sec:low_energy}, the bare momentum-dependent
-interactions in \eqref{eq:H_2} yield only the effective two-body
-Hamiltonian
+operator $\hat k_z^2$ is represented by symmetrized left- and
+right-acting derivative operators in order to preserve manifest
+Hermiticity of $H_{\t{int}}'$.  At third order in the low-energy
+effective theory developed in section \ref{sec:low_energy}, the bare
+momentum-dependent interactions in \eqref{eq:H_2} yield only the
+effective two-body Hamiltonian
 \begin{align}

### 2019-09-13 soc_squeezing/collective_paper/noise_operators.tex
last pre-submission changes to noise operator section

@@
 time-dependent coefficients $\O_{\v m}\p{t}$ for time-independent
-(Schr\"odinger) operators $\hat\S_{\v m}$.  Our discussion clarify why
-noise operators play no role in our calculation of expectation values
-of the form $\bk{\O\p{t}}$ and $\bk{\O\Q\p{t}}$, despite the fact that
-noise operators generally {\it do} need to be considered in the
-calculation of more general multi-time correlators in open quantum
-systems\cite{blocher2019quantum}.
+Schr\"odinger operators $\hat\S_{\v m}$.  Our discussion should
+clarify why noise operators play no role in our calculation of
+expectation values of the form $\bk{\O\p{t}}$ and $\bk{\O\Q\p{t}}$,
+despite the fact that noise operators generally {\it do} need to be
+considered in the calculation of more general multi-time correlators
+in open quantum systems\cite{blocher2019quantum}.
 
@@
 \end{align}
-where we define the operator
+where $\U_{\ul S}^\dag\p{t}$ is the adjoint map of $\U_{\ul S}\p{t}$
+(with respect to a trace inner product between operators on $S$), and
+we define the time-dependent operator
 \begin{align}
@@
 \end{align}
-with $\rho_E$ a time-independent steady state of the environment.
-Enforcing \eqref{eq:markov} for all states $\rho_S$ implies that
+with $\rho_E$ a time-independent steady state of the environment.  If
+we enforce \eqref{eq:markov} for all states $\rho_S$, e.g.~the
+maximally mixed state $\rho_S^{(1)}\propto\1_S$ and
+$\rho_S^{(2)}\equiv\rho_S^{(1)}+\O_S$ with $\O_S$ any traceless
+operator on $S$ with squared norm
+$\norm{\O_S}^2\equiv\tr\p{\O_S^\dag\O_S}\le\tr\1_S$, then by linearity
+we find that
+\begin{align}
+  U_{SE}\p{t} \p{\1_S\otimes\rho_E} U_{SE}^\dag\p{t}
+  \approx \U_{\ul S}\p{t} \1_S\otimes\rho_E,
+  &&
+  U_{SE}\p{t} \p{\O_S\otimes\rho_E} U_{SE}^\dag\p{t}
+  \approx \U_{\ul S}\p{t} \O_S\otimes\rho_E,
+\end{align}
+which implies that the Markov approximation \eqref{eq:markov} holds
+even if we replace $\rho_S$ by any operator on $S$, and in particular
 \begin{align}
   U_{SE}\p{t} \rho_{SE} \p{\O_S\otimes\1_E} U_{SE}^\dag\p{t}
+  = U_{SE}\p{t} \p{\rho_S\O_S\otimes\rho_E} U_{SE}^\dag\p{t}
   \approx \U_{\ul S}\p{t}\p{\rho_S\O_S}\otimes\rho_E.

### 2019-09-13 soc_squeezing/collective_paper/collective_spin.tex
pushing submitted version of collective spin paper

@@
 \newcommand{\sref}[1]{\protect\subref{#1}} % for referencing subfigures
-\graphicspath{{./figures/}} % set path for all figures
 
@@
 molecular, and optical (AMO) systems have enabled the realization of
-collective spin models in a broad range of platforms, including atomic
-ensembles\cite{takano2009spin, appel2009mesoscopic,
-  schleier-smith2010states, chen2011conditional}, Bose-Einstein
-condensates\cite{esteve2008squeezing, riedel2010atomchipbased,
-  gross2010nonlinear, klinder2015dynamical}, ultracold Fermi
-gasses\cite{martin2013quantum}, trapped ions\cite{bohnet2016quantum},
-and optical cavities\cite{leroux2010implementation, bohnet2014reduced,
-  cox2016deterministic, hosten2016measurement,
-  norcia2018cavitymediated}, among others.  These implementations
-compliment theoretical studies in a variety of rich subjects,
-including quantum phase transitions and
+collective spin models in a broad range of platforms, including cold
+atomic gasses\cite{takano2009spin, appel2009mesoscopic}, Bose-Einstein
+condensates\cite{klinder2015dynamical, esteve2008squeezing,
+  riedel2010atomchipbased, gross2010nonlinear}, ultracold Fermi
+gasses\cite{martin2013quantum, bromley2018dynamics,
+  smale2019observation}, trapped ions\cite{bohnet2016quantum}, and
+optical cavities\cite{schleier-smith2010states, chen2011conditional,
+  baumann2010dicke, leroux2010implementation, bohnet2014reduced,
+  cox2016deterministic, hosten2016measurement, hosten2016quantum,
+  norcia2018cavitymediated, ritsch2013cold}, among others.  These
+implementations compliment innumerable theoretical studies in a
+variety of rich subjects, including quantum phase transitions and
 criticality\cite{latorre2005entanglement, alcalde2007functional,
@@
   huang2015twoaxis, muessel2015twistandturn, huang2015quantum,
-  mirkhalaf2018robustifying, lewis-swan2018robust, he2019engineering}.
+  hu2017vacuum, mirkhalaf2018robustifying, lewis-swan2018robust,
+  he2019engineering}.
 
@@
 phonon-mediated\cite{bohnet2016quantum},
-photon-mediated\cite{leroux2010implementation, bohnet2014reduced,
-  cox2016deterministic, hosten2016measurement,
-  norcia2018cavitymediated} or collisional\cite{esteve2008squeezing,
-  riedel2010atomchipbased, gross2010nonlinear, bromley2018dynamics}
-interactions.  Although a truly collective spin model requires
-uniform, all-to-all interactions, as long as measurements do not
-distinguish between constituent particles, even non-uniform systems
-may be effectively described by a uniform model with renormalized
-parameters\cite{hu2015entangled}.
+photon-mediated\cite{klinder2015dynamical, baumann2010dicke,
+  leroux2010implementation, bohnet2014reduced, cox2016deterministic,
+  hosten2016measurement, hosten2016quantum, norcia2018cavitymediated,
+  ritsch2013cold} or collisional\cite{esteve2008squeezing,
+  riedel2010atomchipbased, gross2010nonlinear, martin2013quantum,
+  bromley2018dynamics, smale2019observation} interactions.  Although a
+truly collective spin model requires uniform, all-to-all interactions,
+as long as measurements do not distinguish between constituent
+particles, even non-uniform systems may be effectively described by a
+uniform model with renormalized parameters\cite{hu2015entangled}.
 
@@
 analytical and numerical studies of collective spin systems in the
-future.  We benchmark our method against both exact and quantum
-trajectory Monte Carlo computations of spin squeezing in accessible
-parameter regimes, highlighting both advantages and limitations of our
-short-time expansion.  Finally, we showcase applications of our method
-by computing quantities that are inaccessible to other numerical
-methods.
+future.  We benchmark our method against exact results from both
+analytical calculations and quantum trajectory Monte Carlo
+computations of spin squeezing in accessible parameter regimes,
+highlighting both advantages and limitations of the short-time
+expansion.  Finally, we showcase applications of our method by
+computing quantities that are inaccessible to other numerical methods.
 
@@
 
-\input{calculations_main.tex}
-
-\input{tutorial.tex}
-
-\input{calculations_OAT.tex}
-
-\input{noise_operators.tex}
-
-\input{linear_response.tex}
-
-% \input{changing_bases.tex}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Basic spin operator identities}
+\label{sec:identities}
+
+The appendices in this work make ubiquitous use of various spin
+operator identities; we collect and derive some basic identities here
+for reference.  Note that despite the working definition of collective
+spin operators from $S_\a=\sum_js_\a^{(j)}$, the identities we will
+derive involving only collective spin operators apply just as well to
+large-spin operators that cannot be expressed as the sum of individual
+spin-1/2 operators.  The elementary commutation relations between spin
+operators are, with $\bmu\equiv-\mu\in\set{+1,-1}$ for brevity,
+\begin{align}
+  \sp{s_\z^{(j)},s_\mu^{(k)}}_-
+  &= \delta_{jk} \mu s_\mu^{(j)},
+  &
+  \sp{S_\z,s_\mu^{(j)}}_-
+  &= \sp{s_\z^{(j)},S_\mu}_- = \mu s_\mu^{(j)},
+  &
+  \sp{S_\z,S_\mu}_-
+  &= \mu S_\mu,
+  \label{eq:comm_z_base} \\
+  \sp{s_\mu^{(j)},s_\bmu^{(k)}}_-
+  &= \delta_{jk} 2 \mu s_\z^{(j)},
+  &
+  \sp{S_\mu,s_\bmu^{(j)}}_-
+  &= \sp{s_\mu^{(j)},S_\bmu}_- = 2 \mu s_\z^{(j)},
+  &
+  \sp{S_\mu,S_\bmu}_-
+  &= 2 \mu S_\z.
+  \label{eq:comm_mu_base}
+\end{align}
+These relations can be used to inductively compute identities
+involving powers of collective spin operators.  By pushing through one
+spin operator at a time, we can find
+\begin{align}
+  \p{\mu S_\z}^m s_\mu^{(j)}
+  = \p{\mu S_\z}^{m-1} s_\mu^{(j)} \p{1 + \mu S_\z}
+  = \p{\mu S_\z}^{m-2} s_\mu^{(j)} \p{1 + \mu S_\z}^2
+  = \cdots
+  = s_\mu^{(j)} \p{1 + \mu S_\z}^m,
+  \label{eq:push_z_mu_Ss}
+\end{align}
+and
+\begin{align}
+  \mu s_\z^{(j)} S_\mu^m
+  = S_\mu \mu s_\z^{(j)} S_\mu^{m-1} + s_\mu^{(j)} S_\mu^{m-1}
+  = \cdots
+  = S_\mu^m \mu s_\z^{(j)} + ms_\mu^{(j)} S_\mu^{m-1},
+  \label{eq:push_z_mu_sS}
+\end{align}
+where we will generally find it nicer to express results in terms of
+$\mu s_\z^{(j)}$ and $\mu S_\z$ rather than $s_\z^{(j)}$ and $S_\z$.
+Summing over the single-spin index $j$ in both of the cases above
+gives us the purely collective-spin versions of these identities:
+\begin{align}
+  \p{\mu S_\z}^m S_\mu = S_\mu \p{1 + \mu S_\z}^m,
+  &&
+  \mu S_\z S_\mu^m = S_\mu^m \p{m + \mu S_\z},
+  \label{eq:push_z_mu_single}
+\end{align}
+where we can repeat the process of pushing through individual $S_\z$
+operators $\ell$ times to get
+\begin{align}
+  \p{\mu S_\z}^\ell S_\mu^m
+  = \p{\mu S_\z}^{\ell-1} S_\mu^m \p{m + \mu S_\z}
+  = \p{\mu S_\z}^{\ell-2} S_\mu^m \p{m + \mu S_\z}^2
+  = \cdots
+  = S_\mu^m \p{m + \mu S_\z}^\ell.
+  \label{eq:push_z_mu}
+\end{align}
+Multiplying \eqref{eq:push_z_mu} through by $\p{\mu\nu}^\ell$ (for
+$\nu\in\set{+1,-1}$) and taking its Hermitian conjugate, we can say
+that more generally
+\begin{align}
+  \p{\nu S_\z}^\ell S_\mu^m
+  = S_\mu^m \p{\mu\nu m+\nu S_\z}^\ell,
+  &&
+  S_\mu^m \p{\nu S_\z}^\ell
+  = \p{-\mu\nu m+\nu S_\z}^\ell S_\mu^m.
+\end{align}
+Finding commutation relations between powers of transverse spin
+operators, i.e.~$S_\mu$ and $S_\bmu$, turns out to be considerably
+more difficult than the cases we have worked out thus far.  We
+therefore save this work for Appendix \ref{sec:comm_transverse}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Commutation relations between powers of transverse spin
+  operators}
+\label{sec:comm_transverse}
+
+To find commutation relations between powers of transverse collective
+spin operators, we first compute
+\begin{align}
+  S_\mu^m s_\bmu^{(j)}
+  &= S_\mu^{m-1}s_\bmu^{(j)} S_\mu
+  + S_\mu^{m-1} 2\mu s_\z^{(j)} \\
+  &= S_\mu^{m-2} s_\bmu^{(j)} S_\mu^2
+  + S_\mu^{m-2} 2\mu s_\z^{(j)} S_\mu
+  + S_\mu^{m-1} 2\mu s_\z^{(j)} \\
+  &= s_\bmu^{(j)} S_\mu^m
+  + \sum_{k=0}^{m-1} S_\mu^k 2\mu s_\z^{(j)} S_\mu^{m-k-1}
+  \label{eq:push_mu_Ss_start}.
+\end{align}
+While \eqref{eq:push_mu_Ss_start} gives us the commutator
+$\sp{S_\mu^m,s_\bmu^{(j)}}_-$, we would like to enforce an ordering on
+products of spin operators, which will ensure that we only keep track
+of operators that are linearly independent.  We choose (for now) to
+impose an ordering with all $s_\bmu^{(j)}$ operators on the left, and
+all $s_\z^{(j)}$ operators on the right.  Such an ordering will prove
+convenient for the calculations in this section\footnote{In
+  retrospect, it may have been nicer to push all $s_\mu^{(j)}$
+  operators to the right throughout these calculations, due to the
+  enhanced symmetry that expressions would have with respect to
+  Hermitian conjugation.  In any case, we provide the final result of
+  this section in both ordering conventions, and therefore feel no
+  need to reproduce these calculations with a different ordering of
+  spin operators.}.  This choice of ordering compels us to expand
+\begin{align}
+  \sum_{k=0}^{m-1} S_\mu^k 2\mu s_\z^{(j)} S_\mu^{m-k-1}
+  &= \sum_{k=0}^{m-1} S_\mu^k
+  \sp{2\p{m-k-1} s_\mu^{(j)} S_\mu^{m-k-2}
+    + S_\mu^{m-k-1} 2\mu s_\z^{(j)}} \\
+  &= m \p{m-1} s_\mu^{(j)} S_\mu^{m-2}
+  + m S_\mu^{m-1} 2\mu s_\z^{(j)},
+\end{align}
+which implies
+\begin{align}
+  S_\mu^m s_\bmu^{(j)}
+  = s_\bmu^{(j)} S_\mu^m + m \p{m-1} s_\mu^{(j)} S_\mu^{m-2}
+  + m S_\mu^{m-1} 2\mu s_\z^{(j)},
+  \label{eq:push_mu_Ss}
+\end{align}
+and in turn
+\begin{align}
+  S_\mu^m S_\bmu = S_\bmu S_\mu^m
+  + m S_\mu^{m-1} \p{m - 1 + 2\mu S_\z}.
+  \label{eq:push_mu_single}
+\end{align}
+As the next logical step, we take on the task of computing
+\begin{align}
+  S_\mu^m S_\bmu^n
+  = S_\mu^{m-1} S_\bmu^n S_\mu
+  + n \sp{S_\mu^{m-1} S_\bmu^{n-1} \p{1 - n + 2\mu S_\z}}
+  = S_\bmu^n S_\mu^m
+  + n \sum_{k=0}^{m-1} S_\mu^{m-k-1} S_\bmu^{n-1}
+  \p{1 - n + 2\mu S_\z} S_\mu^k,
+\end{align}
+which implies
+\begin{align}
+  \sp{S_\mu^m, S_\bmu^n}_-
+  = C_{mn;\mu}
+  \equiv n \sum_{k=0}^{m-1} S_\mu^{m-k-1} S_\bmu^{n-1}
+  \p{1 - n + 2\mu S_\z} S_\mu^k.
+\end{align}
+We now need rearrange the operators in $C_{mn;\mu}$ into a standard
+order, which means pushing all $S_\z$ operators to the right and, for
+the purposes of this calculation, all $S_\bmu$ operators to the left.
+We begin by pushing $S_\mu^k$ to the left of $S_\z$, which takes
+$2\mu S_\z\to 2\mu S_\z+2k$, and then push $S_\mu^{m-k-1}$ to the
+right of $S_\bmu^{n-1}$, giving us
+\begin{align}
+  C_{mn;\mu}
+  &= n \sum_{k=0}^{m-1}
+  \p{S_\bmu^{n-1} S_\mu^{m-k-1} + C_{m-k-1,n-1;\mu}} S_\mu^k
+  \p{2k + 1 - n + 2\mu S_\z} \\
+  &= D_{mn;\mu}
+  + n \sum_{k=0}^{m-2} C_{m-k-1,n-1;\mu}
+  S_\mu^k \p{2k + 1 - n + 2\mu S_\z},
+  \label{eq:C_mn}
+\end{align}
+where we have dropped the last ($k=m-1$) term in the remaining sum
+because $C_{m-k-1,n-1;\mu}=0$ if $k=m-1$, and
+\begin{align}
+  D_{mn;\mu}
+  \equiv mn S_\bmu^{n-1} S_\mu^{m-1} \p{m - n + 2\mu S_\z}.
+  \label{eq:D_mn}
+\end{align}
+To our despair, we have arrived in \eqref{eq:C_mn} at a {\it
+  recursive} formula for $C_{mn;\mu}$.  Furthermore, we have not even
+managed to order all spin operators, as $C_{m-k-1,n-1;\mu}$ contains
+$S_\z$ operators that are to the left of $S_\mu^k$.  To sort all spin
+operators once and for all, we define
+\begin{align}
+  C_{mn;\mu}^{(k)} \equiv C_{m-k,n;\mu} S_\mu^k,
+  &&
+  D_{mn;\mu}^{(k)} \equiv D_{m-k,n;\mu} S_\mu^k,
+\end{align}
+which we can expand as
+\begin{align}
+  D_{mn;\mu}^{(k)}
+  &= \p{m-k}n S_\bmu^{n-1} S_\mu^{m-k-1}
+  \p{m-k-n+2\mu S_\z} S_\mu^k \\
+  &= \p{m-k}n S_\bmu^{n-1} S_\mu^{m-1} \p{k+m-n+2\mu S_\z},
+  \label{eq:D_mn_k}
+\end{align}
+and
+\begin{align}
+  C_{mn;\mu}^{(k)}
+  &= D_{m-k,n;\mu} S_\mu^k + n \sum_{j=0}^{m-k-2}
+  C_{m-k-j-1,n-1;\mu} S_\mu^j \p{2j+1-n+2\mu S_\z} S_\mu^k \\
+  &= D_{mn;\mu}^{(k)} + n \sum_{j=0}^{m-k-2}
+  C_{m-k-j-1,n-1;\mu} S_\mu^{j+k} \p{2j+2k+1-n+2\mu S_\z} \\
+  &= D_{mn;\mu}^{(k)} + n \sum_{j=0}^{m-k-2}
+  C_{m-1,n-1;\mu}^{(k+j)} \p{2\sp{j+k}+1-n+2\mu S_\z} \\
+  &= D_{mn;\mu}^{(k)} + n \sum_{j=k}^{m-2}
+  C_{m-1,n-1;\mu}^{(j)} \p{2j+1-n+2\mu S_\z}.
+  \label{eq:C_mn_k}
+\end{align}
+While the resulting expression in \eqref{eq:C_mn_k} strongly resembles
+that in \eqref{eq:C_mn}, there is one crucial difference: all spin
+operators in \eqref{eq:C_mn_k} have been sorted into a standard order.
+We can now repeatedly substitute $C_{mn;\mu}^{(k)}$ into itself, each
+time decreasing $m$ and $n$ by 1, until one of $m$ or $n$ reaches
+zero.  Such repeated substitution yields the expansion
+\begin{align}
+  C_{mn;\mu}
+  = C_{mn;\mu}^{(0)}
+  = D_{mn;\mu}
+  + \sum_{p=1}^{\min\set{m,n}-1} E_{mn;\mu}^{(p)},
+  \label{eq:C_mn_E}
+\end{align}
+where the first two terms in the sum over $p$ are
+\begin{align}
+  E_{mn;\mu}^{(1)}
+  &= n \sum_{k=0}^{m-2} D_{m-1,n-1;\mu}^{(k)} \p{2k+1-n+2\mu S_\z}, \\
+  E_{mn;\mu}^{(2)}
+  &= n \sum_{k_1=0}^{m-2} \p{n-1} \sum_{k_2=k_1}^{m-3}
+  D_{m-2,n-2;\mu}^{(k_2)} \p{2k_2+2-n+2\mu S_\z} \p{2k_1+1-n+2\mu S_\z},
+\end{align}
+and more generally for $p>1$,
+\begin{align}
+  E_{mn;\mu}^{(p)}
+  = \f{n!}{\p{n-p}!}
+  \sum_{k_1=0}^{m-2} \sum_{k_2=k_1}^{m-3} \cdots\sum_{k_p=k_{p-1}}^{m-p-1}
+  D_{m-p,n-p;\mu}^{(k_p)} \prod_{j=1}^p \p{2k_j+j-n+2\mu S_\z}.
+  \label{eq:E_mn_p}
+\end{align}
+In principle, the expressions in \eqref{eq:D_mn}, \eqref{eq:D_mn_k},
+\eqref{eq:C_mn_E}, and \eqref{eq:E_mn_p} suffice to evaluate the
+commutator $\sp{S_\mu^m,S_\bmu^n}_- = C_{mn;\mu}$, but this result is
+-- put lightly -- quite a mess: the expression for $E_{mn;\mu}^{(p)}$
+in \eqref{eq:E_mn_p} involves a sum over $p$ mutually dependent
+intermediate variables, each term of which additionally contains a
+product of $p$ factors.  We therefore devote the rest of this section
+to simplifying our result for the commutator
+$\sp{S_\mu^m,S_\bmu^n}_-$.
+
+Observing that in \eqref{eq:E_mn_p} we always have
+$0\le k_1\le k_2\le\cdots\le k_p\le m-p-1$, we can rearrange the order
+of the sums and relabel $k_p\to\ell$ to get
+\begin{align}
+  E_{mn;\mu}^{(p)}
+  = \f{n!}{\p{n-p}!}
+  \sum_{\ell=0}^{m-p-1} D_{m-p,n-p;\mu}^{(\ell)} \p{2\ell+F_{np;\mu}}
+  \sum_{\p{\v k,p-1,\ell}} \prod_{j=1}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}},
+  \label{eq:E_mn_p_sum}
+\end{align}
+where for shorthand we define
+\begin{align}
+  F_{np;\mu} \equiv p - n + 2\mu S_\z,
+  &&
+  \sum_{\p{\v k,q,\ell}} X \equiv
+  \sum_{k_1=0}^\ell \sum_{k_2=k_1}^\ell
+  \cdots \sum_{k_q=k_{q-1}}^\ell X.
+\end{align}
+We now further define
+\begin{align}
+  f_{np\ell;\mu}\p{k,q} \equiv \p{\ell-k+q} \p{\ell+k-q+F_{np;\mu}},
+\end{align}
+and evaluate sums successively over $k_{p-1},k_{p-2},\cdots,k_1$,
+finding
+\begin{align}
+  \sum_{\p{\v k,p-1,\ell}} \prod_{j=1}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}}
+  &= \sum_{\p{\v k,p-2,\ell}}
+  \prod_{j=2}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}}
+  f_{np\ell;\mu}\p{k_{p-2},1} \\
+  &= \f1{\p{r-1}!} \sum_{\p{\v k,p-r,\ell}}
+  \prod_{j=r}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}}
+  \prod_{q=1}^{r-1} f_{np\ell;\mu}\p{k_{p-r},q} \\
+  &= \f1{\p{p-1}!} \prod_{q=1}^{p-1} f_{np\ell;\mu}\p{0,q} \\
+  &= { \ell + p - 1 \choose p - 1 }
+  \prod_{q=1}^{p-1} \p{\ell-q+F_{np;\mu}}.
+\end{align}
+Substitution of this result together with $D_{m-p,n-p;\mu}^{(\ell)}$
+using \eqref{eq:D_mn_k} into \eqref{eq:E_mn_p_sum} then gives us
+\begin{align}
+  E_{mn;\mu}^{(p)}
+  = \f{n!}{\p{n-p-1}!} S_\bmu^{n-p-1} S_\mu^{m-p-1} G_{mnp;\mu}
+\end{align}
+with
+\begin{align}
+  G_{mnp;\mu}
+  &\equiv \sum_{\ell=0}^{m-p-1} { \ell + p - 1 \choose p - 1 }
+  \p{m-p-\ell} \p{\ell+m-p+F_{np;\mu}}
+  \p{2\ell + F_{np;\mu}}
+  \prod_{q=1}^{p-1} \p{\ell-q+F_{np;\mu}} \\
+  &= { m \choose p + 1 } \prod_{q=0}^p \p{m-p-q+F_{np;\mu}}.
+\end{align}
+We can further simplify
+\begin{align}
+  \prod_{q=0}^p \p{m-p-q+F_{np;\mu}}
+  = \prod_{q=0}^p \p{m-n-q+2\mu S_\z}
+  = \sum_{q=0}^{p+1} \p{-1}^{p+1-q}
+  { p+1 \brack q } \p{m-n+2\mu S_\z}^q,
+\end{align}
+where ${ p \brack q }$ is an unsigned Stirling number of the first
+kind, and finally
+\begin{align}
+  \sum_{q=0}^p \p{-1}^{p-q} { p \brack q } \p{m-n+2\mu S_\z}^q
+  &= \sum_{q=0}^p \p{-1}^{p-q} { p \brack q } \sum_{\ell=0}^q
+  { q \choose \ell } \p{m-n}^{q-\ell} \p{2\mu S_\z}^\ell \\
+  &= \sum_{\ell=0}^p 2^\ell \sum_{q=\ell}^p \p{-1}^{p-q}
+  { p \brack q } { q \choose \ell } \p{m-n}^{q-\ell} \p{\mu S_\z}^\ell.
+\end{align}
+Putting everything together, we finally have
+\begin{align}
+  E_{mn;\mu}^{(p-1)}
+  = p! { m \choose p } { n \choose p }
+  S_\bmu^{n-p} S_\mu^{m-p}
+  \sum_{\ell=0}^p \epsilon_{mn}^{p\ell} \p{\mu S_\z}^\ell,
+\end{align}
+with
+\begin{align}
+  \epsilon_{mn}^{p\ell}
+  \equiv 2^\ell \sum_{q=\ell}^p \p{-1}^{p-q}
+  { p \brack q } { q \choose \ell } \p{m-n}^{q-\ell},
+\end{align}
+where in this final form $E_{mn;\mu}^{(0)} = D_{mn;\mu}$, which
+together with the expansion for $C_{mn;\mu}$ in \eqref{eq:C_mn_E}
+implies that
+\begin{align}
+  \sp{S_\mu^m, S_\bmu^n}_-
+  = \sum_{p=1}^{\min\set{m,n}}
+  p! { m \choose p } { n \choose p } S_\bmu^{n-p} S_\mu^{m-p}
+  \sum_{\ell=0}^p \epsilon_{mn}^{p\ell} \p{\mu S_\z}^\ell,
+  \label{eq:comm_mu}
+\end{align}
+and
+\begin{align}
+  S_\mu^m S_\bmu^n
+  = \sum_{p=0}^{\min\set{m,n}}
+  p! { m \choose p } { n \choose p } S_\bmu^{n-p} S_\mu^{m-p}
+  \sum_{\ell=0}^p \epsilon_{mn}^{p\ell} \p{\mu S_\z}^\ell.
+  \label{eq:push_mu_bmu}
+\end{align}
+If we wish to order products of collective spin operators with $S_\z$
+in between $S_\bmu$ and $S_\mu$, then
+\begin{align}
+  S_\mu^m S_\bmu^n
+  = \sum_{p=0}^{\min\set{m,n}} p! { m \choose p } { n \choose p }
+  S_\bmu^{n-p} Z_{mn;\bmu}^{(p)} S_\mu^{m-p},
+\end{align}
+where
+\begin{align}
+  Z_{mn;\bmu}^{(p)}
+  \equiv \sum_{\ell=0}^p \epsilon_{mn}^{p\ell}
+  \p{-\sp{m-p} + \mu S_\z}^\ell
+  = \sum_{q=0}^p \zeta_{mn}^{pq} \p{\bmu S_\z}^q,
+  \label{eq:Z_mnp}
+\end{align}
+with
+\begin{align}
+  \zeta_{mn}^{pq}
+  \equiv \sum_{\ell=q}^p \epsilon_{mn}^{p\ell}
+  { \ell \choose q } \p{-1}^\ell \p{m-p}^{\ell-q}
+  = \p{-1}^p 2^q \sum_{s=q}^p
+  { p \brack s } { s \choose q } \p{m+n-2p}^{s-q}.
+  \label{eq:zeta_mnpq}
+\end{align}
+Here ${ p \brack s }$ is an unsigned Stirling number of the first
+kind.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Product of arbitrary ordered collective spin operators}
+\label{sec:general_product}
+
+The most general product of collective spin operators that we need to
+compute is
+\begin{align}
+  \S^{pqr}_{\ell mn;\mu}
+  = S_\mu^p \p{\mu S_\z}^q S_\bmu^r
+  S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n
+  = \sum_{k=0}^{\min\set{r,\ell}} k! { r \choose k } { \ell \choose k }
+  S_\mu^{p+\ell-k} \tilde Z_{qr\ell m;\mu}^{(k)} S_\bmu^{r+n-k},
+  \label{eq:general_product}
+\end{align}
+where
+\begin{align}
+  \tilde Z_{qr\ell m;\mu}^{(k)}
+  &\equiv \p{\ell-k+\mu S_\z}^q
+  Z_{r\ell;\mu}^{(k)} \p{r-k+\mu S_\z}^m \\
+  &= \sum_{a=0}^k \zeta_{r\ell}^{ka}
+  \sum_{b=0}^q \p{\ell-k}^{q-b} { q \choose b }
+  \sum_{c=0}^m \p{r-k}^{m-c} { m \choose c }
+  \p{\mu S_\z}^{a+b+c},
+\end{align}
+is defined in terms of $Z_{r\ell;\mu}^{(k)}$ and $\zeta_{r\ell}^{ka}$
+as respectively given in \eqref{eq:Z_mnp} and \eqref{eq:zeta_mnpq}.
+The (anti-)commutator of two ordered products of collective spin
+operators is then simply
+\begin{align}
+  \sp{S_\mu^p \p{\mu S_\z}^q S_\bmu^r,
+    S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_\pm
+  = \S^{pqr}_{\ell mn;\mu} \pm \S^{\ell mn}_{pqr;\mu}.
+\end{align}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Sandwich identities for single-spin decoherence calculations}
+\label{sec:sandwich_single}
+
+In this section we derive several identities that will be necessary
+for computing the effects of single-spin decoherence on ordered
+products of collective spin operators, i.e.~on operators of the form
+$S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$.  These identities all involve
+sandwiching a collective spin operator between operators that act on
+individual spins only, and summing over all individual spin indices.
+Our general strategy will be to use commutation relations to push
+single-spin operators together, and then evaluate the sum to arrive at
+an expression involving only collective spin operators.
+
+We first compute sums of single-spin operators sandwiching
+$\p{\mu S_\z}^m$, when necessary making use of the identity in
+\eqref{eq:push_z_mu_Ss}.  The unique cases up to Hermitian conjugation
+are, for $S\equiv N/2$ and $\mu,\nu\in\set{+1,-1}$,
+\begin{align}
+  \sum_j s_\z^{(j)} \p{\mu S_\z}^m s_\z^{(j)}
+  &= \sum_j s_\z^{(j)} s_\z^{(j)} \p{\mu S_\z}^m
+  = \f14 \sum_j \1_j \p{\mu S_\z}^m
+  = \f12 S \p{\mu S_\z}^m, \\
+  \sum_j s_\z^{(j)} \p{\mu S_\z}^m s_\nu^{(j)}
+  &= \p{\mu S_\z}^m \sum_j s_\z^{(j)} s_\nu^{(j)}
+  = \f12 \p{\mu S_\z}^m \nu S_\nu
+  = \f12 \nu S_\nu \p{\mu\nu+\mu S_\z}^m, \\
+  \sum_j s_\nu^{(j)} \p{\mu S_\z}^m s_\nu^{(j)}
+  &= \sum_j s_\nu^{(j)} s_\nu^{(j)} \p{\mu\nu+\mu S_\z}^m
+  = 0, \\
+  \sum_j s_\bnu^{(j)} \p{\mu S_\z}^m s_\nu^{(j)}
+  &= \sum_j s_\bnu^{(j)} s_\nu^{(j)} \p{\mu\nu+\mu S_\z}^m
+  = \p{S-\nu S_\z} \p{\mu\nu+\mu S_\z}^m.
+\end{align}
+We are now equipped to derive similar identities for more general
+collective spin operators.  Making heavy use of identities
+\eqref{eq:push_z_mu_sS} and \eqref{eq:push_mu_Ss} to push single-spin
+operators through transverse collective-spin operators, we again work
+through all combinations that are unique up to Hermitian conjugation,
+finding
+\begin{align}
+  \sum_j s_\z^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\z^{(j)}
+  &= \f12 \p{S-\ell-n} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n
+  + \ell n S_\mu^{\ell-1} \p{S+\mu S_\z}
+  \p{-1+\mu S_\z}^m S_\bmu^{n-1},
+  \label{eq:san_z_z} \allowdisplaybreaks \\
+  \sum_j s_\z^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\mu^{(j)}
+  &= \f12 \mu S_\mu^{\ell+1} \p{1+\mu S_\z}^m S_\bmu^n
+  - \mu n \p{S-\ell-\f12\sp{n-1}} S_\mu^\ell
+  \p{\mu S_\z}^m S_\bmu^{n-1} \notag \\
+  &\qquad - \mu\ell n\p{n-1} S_\mu^{\ell-1}
+  \p{S+\mu S_\z} \p{-1+\mu S_\z}^m S_\bmu^{n-2},
+  \label{eq:san_z_mu} \allowdisplaybreaks \\
+  \sum_j s_\z^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\bmu^{(j)}
+  &= -\f12 \mu S_\mu^\ell \p{\mu S_\z}^m S_\bmu^{n+1}
+  + \mu \ell S_\mu^{\ell-1} \p{S+\mu S_\z} \p{-1+\mu S_\z}^m S_\bmu^n,
+  \label{eq:san_z_bmu} \allowdisplaybreaks \\
+  \sum_j s_\mu^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\mu^{(j)}
+  &= n S_\mu^{\ell+1} \p{\mu S_\z}^m S_\bmu^{n-1}
+  - n\p{n-1} S_\mu^\ell \p{S+\mu S_\z} \p{-1+\mu S_\z}^m S_\bmu^{n-2},
+  \label{eq:san_mu_mu} \allowdisplaybreaks \\
+  \sum_j s_\mu^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\bmu^{(j)}
+  &= S_\mu^\ell \p{S+\mu S_\z}\p{-1+\mu S_\z}^m S_\bmu^n,
+  \label{eq:san_mu_bmu} \allowdisplaybreaks \\
+  \sum_j s_\bmu^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\mu^{(j)}
+  &= S_\mu^\ell \p{S - \ell - n - \mu S_\z}
+  \p{1+\mu S_\z}^m S_\bmu^n
+  + \ell n \p{2S - \ell - n + 2}
+  S_\mu^{\ell-1} \p{\mu S_\z}^m S_\bmu^{n-1} \notag \\
+  &\qquad + \ell n \p{\ell-1} \p{n-1} S_\mu^{\ell-2} \p{S+\mu S_\z}
+  \p{-1+\mu S_\z}^m S_\bmu^{n-2}.
+  \label{eq:san_bmu_mu}
+\end{align}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Uncorrelated, permutationally-symmetric single-spin
+  decoherence}
+\label{sec:decoherence_single}
+
+In this section we work out the effects of permutationally-symmetric
+decoherence of individual spins on collective spin operators of the
+form $S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$.  For compactness, we define
+\begin{align}
+  \D\p{g} \O
+  \equiv \D\p{\set{g^{(j)}:j=1,2,\cdots,N}} \O
+  = \sum_j\p{{g^{(j)}}^\dag \O g^{(j)}
+    - \f12\sp{{g^{(j)}}^\dag g^{(j)}, \O}_+},
+\end{align}
+where $g$ is an operator that acts on a single spin, $g^{(j)}$ is an
+operator that acts with $g$ on spin $j$ and trivially on all other
+spins, and $N$ is the total number of spins.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Decay-type decoherence}
+\label{sec:decay_single}
+
+The effect of decoherence via uncorrelated decay ($\mu=-1$) or
+excitation ($\mu=1$) of individual spins is described by
+\begin{align}
+  \D\p{s_\mu} \O
+  = \sum_j\p{s_\bmu^{(j)} \O s_\mu^{(j)}
+    - \f12\sp{s_\bmu^{(j)} s_\mu^{(j)},\O}_+}
+  = \sum_j s_\bmu^{(j)} \O s_\mu^{(j)}
+  - S \O + \f{\mu}{2} \sp{S_\z, \O}_+.
+\end{align}
+In order to determine the effect of this decoherence on general
+collective spin operators, we expand the anti-commutator
+\begin{align}
+  \sp{S_\z, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
+  = S_\z S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n
+  + S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\z
+  = \mu S_\mu^\ell\p{\ell+n+2\mu S_\z} \p{\mu S_\z}^m S_\bmu^n,
+\end{align}
+which implies, using \eqref{eq:san_mu_bmu},
+\begin{align}
+  \D\p{s_\bmu} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
+  = S_\mu^\ell \p{S+\mu S_\z}\p{-1+\mu S_\z}^m S_\bmu^n
+  - S_\mu^\ell\sp{S + \f12\p{\ell+n} + \mu S_\z}
+  \p{\mu S_\z}^m S_\bmu^n,
+  \label{eq:decay_diff}
+\end{align}
+and, using \eqref{eq:san_bmu_mu},
+\begin{align}
+  \D\p{s_\mu} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
+  &= S_\mu^\ell \p{S - \ell - n - \mu S_\z} \p{1+\mu S_\z}^m S_\bmu^n
+  - S_\mu^\ell\sp{S - \f12\p{\ell+n} - \mu S_\z}
+  \p{\mu S_\z}^m S_\bmu^n \notag \\
+  &\qquad + \ell n \p{2S - \ell - n + 2}
+  S_\mu^{\ell-1} \p{\mu S_\z}^m S_\bmu^{n-1} \notag \\
+  &\qquad + \ell n \p{\ell-1} \p{n-1} S_\mu^{\ell-2} \p{S + \mu S_\z}
+  \p{-1+\mu S_\z}^m S_\bmu^{n-2}.
+  \label{eq:decay_same}
+\end{align}
+Decoherence via jump operators $s_\bmu^{(j)}$ only couples operators
+$S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$ to operators
+$S_\mu^\ell \p{\mu S_\z}^{m'} S_\bmu^n$ with $m'\le m$.  Decoherence
+via jump operators $s_\mu^{(j)}$, meanwhile, makes operators
+$S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$ ``grow'' in $m$ through the last
+term in \eqref{eq:decay_same}, although the sum $\ell+m+n$ does not
+grow.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Dephasing}
+\label{sec:dephasing_single}
+
+The effect of decoherence via single-spin dephasing is described by
+\begin{align}
+  \D\p{s_\z} \O
+  = \sum_j\p{s_\z^{(j)} \O s_\z^{(j)}
+    - \f12\sp{s_\z^{(j)} s_\z^{(j)},\O}_+}
+  = \sum_j s_\z^{(j)} \O s_\z^{(j)} - \f12 S \O.
+\end{align}
+From \eqref{eq:san_z_z}, we then have
+\begin{align}
+  \D\p{s_\z} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
+  = -\f12\p{\ell+n} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n
+  + \ell n S_\mu^{\ell-1} \p{S + \mu S_\z}
+  \p{-1 + \mu S_\z}^m S_\bmu^{n-1}.
+\end{align}
+Decoherence via single-spin dephasing makes operators
+$S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$ ``grow'' in $m$, although the sum
+$\ell+m+n$ does not grow.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{The general case}
+\label{sec:general_single}
+
+The most general type of single-spin decoherence is described by
+\begin{align}
+  \D\p{g} \O
+  = \sum_j\p{{g^{(j)}}^\dag \O g^{(j)}
+    - \f12\sp{{g^{(j)}}^\dag g^{(j)}, \O}_+},
+  &&
+  g \equiv g_\z s_\z + g_+ s_+ + g_- s_-.
+  \label{eq:D_general_single}
+\end{align}
+To simplify \eqref{eq:D_general_single}, we expand
+\begin{align}
+  g^\dag \O g
+  = \abs{g_\z}^2 s_\z \O s_\z
+  + \sum_\mu \p{\abs{g_\mu}^2 s_\bmu \O s_\mu
+    + g_\bmu^* g_\mu s_\mu \O s_\mu
+    + g_\z^* g_\mu s_\z \O s_\mu
+    + g_\bmu^* g_\z s_\mu \O s_\z},
+\end{align}
+and
+\begin{align}
+  g^\dag g
+  = \f14 \abs{g_\z}^2
+  + \f12 \sum_\mu \sp{\abs{g_\mu}^2 \p{1-2\mu s_\z}
+    + \mu \p{g_\z^*g_\mu - g_\bmu^*g_\z} s_\mu},
+\end{align}
+which implies
+\begin{align}
+  \D\p{g} \O
+  &= \sum_{X\in\set{\z,+,-}} \abs{g_X}^2 \D\p{s_X} \O
+  + \sum_{\mu,j}
+  \p{g_\bmu^* g_\mu s_\mu^{(j)} \O s_\mu^{(j)}
+    + g_\z^* g_\mu s_\z^{(j)} \O s_\mu^{(j)}
+    + g_\bmu^* g_\z s_\mu^{(j)} \O s_\z^{(j)}}
+  \notag \\
+  &\qquad -\f14 \sum_\mu \mu
+  \p{g_\z^*g_\mu - g_\bmu^*g_\z} \sp{S_\mu, \O}_+.
+\end{align}
+In order to compute the effect of this decoherence on general
+collective spin operators, we expand the anti-commutator
+\begin{align}
+  \sp{S_\mu, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
+  = S_\mu^{\ell+1} \sp{\p{\mu S_\z}^m+\p{1+\mu S_\z}^m} S_\bmu^n
+  - n S_\mu^\ell \p{n-1+2\mu S_\z} \p{\mu S_\z}^m S_\bmu^{n-1}.
+  \label{eq:S_mu_acomm}
+\end{align}
+Recognizing a resemblance between terms in \eqref{eq:S_mu_acomm} and
+\eqref{eq:san_z_mu}, we collect terms to simplify
+\begin{align}
+  \sum_j s_\z^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\mu^{(j)}
+  - \f14 \mu \sp{S_\mu, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
+  = K_{\ell mn;\mu} + L_{\ell mn;\mu}
+  \label{eq:dec_z_mu}
+\end{align}
+and likewise
+\begin{align}
+  \sum_j s_\mu^{(j)} S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n s_\z^{(j)}
+  + \f14 \mu \sp{S_\mu, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
+  = K_{\ell mn;\mu} + M_{\ell mn;\mu}
+  \label{eq:dec_mu_z}
+\end{align}
+with
+\begin{align}
+  K_{\ell mn;\mu}
+  &\equiv \f14 \mu S_\mu^{\ell+1}
+  \sp{\p{1+\mu S_\z}^m-\p{\mu S_\z}^m} S_\bmu^n, \\
+  L_{\ell mn;\mu}
+  &\equiv -\mu n S_\mu^\ell \sp{S-\ell-\f34\p{n-1}-\f12\mu S_\z}
+  \p{\mu S_\z}^m S_\bmu^{n-1}
+  - \mu\ell n\p{n-1} S_\mu^{\ell-1}
+  \p{S+\mu S_\z} \p{-1+\mu S_\z}^m S_\bmu^{n-2}, \\
+  M_{\ell mn;\mu}
+  &\equiv \mu n S_\mu^\ell \sp{\p{S+\mu S_\z}\p{-1+\mu S_\z}^m
+    - \f12\p{\f12\sp{n-1}+\mu S_\z}\p{\mu S_\z}^m} S_\bmu^{n-1}.
+\end{align}
+Defining for completion
+\begin{align}
+  P_{\ell mn;\mu}
+  &\equiv \sum_j s_\mu^{(j)} S_\mu^\ell
+  \p{\mu S_\z}^m S_\bmu^n s_\mu^{(j)}
+  = n S_\mu^{\ell+1} \p{\mu S_\z}^m S_\bmu^{n-1}
+  - n\p{n-1} S_\mu^\ell \p{S+\mu S_\z} \p{-1+\mu S_\z}^m S_\bmu^{n-2},
+\end{align}
+and
+\begin{align}
+  Q_{\ell mn;\mu}^{(g)}
+  \equiv g_\bmu^* g_\mu P_{\ell mn;\mu}
+  + \p{g_\z^* g_\mu + g_\bmu^* g_\z}
+  K_{\ell mn;\mu}
+  + g_\z^* g_\mu L_{\ell mn;\mu}
+  + g_\bmu^* g_\z M_{\ell mn;\mu},
+  \label{eq:Q_single}
+\end{align}
+we finally have
+\begin{align}
+  \D\p{g} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
+  = \sum_{X\in\set{\z,+,-}} \abs{g_X}^2
+  \D\p{s_X} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
+  + Q_{\ell mn;\mu}^{(g)} + \sp{Q_{nm\ell;\mu}^{(g)}}^\dag.
+\end{align}
+Note that the sum $\ell+m+n$ for operators
+$S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$ does not grow under this type of
+decoherence.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Sandwich identities for collective-spin decoherence
+  calculations}
+\label{sec:sandwich_collective}
+
+In analogy with the work in Appendix \ref{sec:sandwich_single}, in
+this section we work out sandwich identities necessary for
+collective-spin decoherence calculations.  The simplest cases are
+\begin{align}
+  S_\mu S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\bmu
+  &= S_\mu^{\ell+1} \p{\mu S_\z}^m S_\bmu^{n+1},
+  \allowdisplaybreaks \\
+  S_\mu S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\z
+  &= \mu S_\mu^{\ell+1} \p{n+\mu S_\z} \p{\mu S_\z}^m S_\bmu^n,
+  \allowdisplaybreaks \\
+  S_\z S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\z
+  &= S_\mu^\ell \sp{\ell n + \p{\ell+n} \mu S_\z + \p{\mu S_\z}^2}
+  \p{\mu S_\z}^m S_\bmu^n.
+\end{align}
+With a bit more work, we can also find
+\begin{align}
+  S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\mu
+  &= S_\mu^{\ell+1} \p{1+\mu S_\z}^m S_\bmu^n
+  - n S_\mu^\ell \p{n-1+2\mu S_\z} \p{\mu S_\z}^m S_\bmu^{n-1},
+\end{align}
+which implies
+\begin{align}
+  S_\mu S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\mu
+  &= S_\mu^{\ell+2} \p{1+\mu S_\z}^m S_\bmu^n
+  - n S_\mu^{\ell+1} \p{n-1+2\mu S_\z} \p{\mu S_\z}^m S_\bmu^{n-1},
+  \allowdisplaybreaks \\
+  S_\z S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\mu
+  &= \mu S_\mu^{\ell+1} \p{\ell+1+\mu S_\z} \p{1+\mu S_\z}^m S_\bmu^n
+  \notag \\
+  &\qquad - \mu n S_\mu^\ell
+  \sp{\ell\p{n-1} + \p{2\ell+n-1}\mu S_\z + 2\p{\mu S_\z}^2}
+  \p{\mu S_\z}^m S_\bmu^{n-1}.
+\end{align}
+Finally, we compute
+\begin{align}
+  S_\bmu S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\mu
+  &= \sp{S_\mu^\ell S_\bmu - \ell S_\mu^{\ell-1} \p{\ell-1+2\mu S_\z}}
+  \p{\mu S_\z}^m
+  \sp{S_\mu S_\bmu^n - n \p{n-1+2\mu S_\z} S_\bmu^{n-1}} \notag \\
+  &= S_\mu^\ell S_\bmu \p{\mu S_\z}^m S_\mu S_\bmu^n \notag \\
+  &\qquad - S_\mu^\ell
+  \sp{\ell\p{\ell+1} + n\p{n+1}+2\p{\ell+n}\mu S_\z}
+  \p{1+\mu S_\z}^m S_\bmu^n \notag \\
+  &\qquad + \ell n S_\mu^{\ell-1}
+  \sp{\p{\ell-1}\p{n-1}+2\p{\ell+n-2}\mu S_\z + 4\p{\mu S_\z}^2}
+  \p{\mu S_\z}^m S_\bmu^{n-1},
+\end{align}
+where
+\begin{align}
+  S_\bmu \p{\mu S_\z}^m S_\mu
+  = S_\bmu S_\mu \p{1+\mu S_\z}^m
+  = \p{S_\mu S_\bmu - 2\mu S_\z} \p{1+\mu S_\z}^m
+  = S_\mu \p{2+\mu S_\z}^m S_\bmu - 2\mu S_\z \p{1+\mu S_\z}^m,
+\end{align}
+so
+\begin{align}
+  S_\bmu S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n S_\mu
+  &= S_\mu^{\ell+1} \p{2+\mu S_\z}^m S_\bmu^{n+1} \notag \\
+  &\qquad - S_\mu^\ell
+  \sp{\ell\p{\ell+1} + n\p{n+1}+2\p{\ell+n+1}\mu S_\z}
+  \p{1+\mu S_\z}^m S_\bmu^n \notag \\
+  &\qquad + \ell n S_\mu^{\ell-1}
+  \sp{\p{\ell-1}\p{n-1}+2\p{\ell+n-2}\mu S_\z + 4\p{\mu S_\z}^2}
+  \p{\mu S_\z}^m S_\bmu^{n-1}.
+\end{align}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Collective spin decoherence}
+\label{sec:decoherence_collective}
+
+In this section we work out the effects of collective decoherence on
+general collective spin operators.  For shorthand, we define
+\begin{align}
+  \D\p{G} \O
+  \equiv \D\p{\set{G}} \O
+  = G^\dag \O G - \f12\sp{G^\dag G, \O}_+,
+\end{align}
+where $G$ is a collective spin jump operator.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Decay-type decoherence and dephasing}
+\label{sec:decay_dephasing_collective}
+
+Making use of the results in Appendix \ref{sec:sandwich_collective},
+we find that the effects of collective decay-type decoherence on
+general collective spin operators are given by
+\begin{align}
+  \D\p{S_\bmu} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
+  &= -S_\mu^{\ell+1} \sp{\p{1+\mu S_\z}^m - \p{\mu S_\z}^m} S_\bmu^{n+1}
+  \notag \\
+  &\qquad + \f12 S_\mu^\ell \sp{\ell\p{\ell-1} + n\p{n-1}
+    + 2\p{\ell+n}\mu S_\z} \p{\mu S_\z}^m S_\bmu^n,
+\end{align}
+and
+\begin{align}
+  \D\p{S_\mu} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
+  &= S_\mu^{\ell+1} \sp{\p{2+\mu S_\z}^m-\p{1+\mu S_\z}^m} S_\bmu^{n+1}
+  \notag \\
+  &\qquad - S_\mu^\ell
+  \sp{\ell\p{\ell+1} + n\p{n+1}+2\p{\ell+n+1}\mu S_\z}
+  \p{1+\mu S_\z}^m S_\bmu^n \notag \\
+  &\qquad + \f12 S_\mu^\ell
+  \sp{\ell\p{\ell+1} + n\p{n+1}+2\p{\ell+n+2}\mu S_\z}
+  \p{\mu S_\z}^m S_\bmu^n \notag \\
+  &\qquad + \ell n S_\mu^{\ell-1}
+  \sp{\p{\ell-1}\p{n-1}+2\p{\ell+n-2}\mu S_\z + 4\p{\mu S_\z}^2}
+  \p{\mu S_\z}^m S_\bmu^{n-1}.
+\end{align}
+Similarly, the effect of collective dephasing is given by
+\begin{align}
+  \D\p{S_\z} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
+  = -\f12 \p{\ell-n}^2 S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n.
+\end{align}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{The general case}
+\label{sec:general_collective}
+
+More generally, we consider jump operators of the form
+\begin{align}
+  G \equiv G_\z S_\z + G_+ S_+ + G_- S_-,
+\end{align}
+whose decoherence effects are determined by
+\begin{align}
+  G^\dag \O G
+  = \abs{G_\z}^2 S_\z \O S_\z
+  + \sum_\mu \p{\abs{G_\mu}^2 S_\bmu \O S_\mu
+    + G_\bmu^* G_\mu S_\mu \O S_\mu
+    + G_\z^* G_\mu S_\z \O S_\mu
+    + G_\bmu^* G_\z S_\mu \O S_\z},
+\end{align}
+and
+\begin{align}
+  G^\dag G
+  = \abs{G_\z}^2 S_\z^2
+  + \sum_\mu \p{\abs{G_\mu}^2 S_\bmu S_\mu
+    + G_\z^*G_\mu S_\z S_\mu
+    + G_\bmu^* G_\z S_\mu S_\z
+    + G_\bmu^* G_\mu S_\mu^2},
+\end{align}
+which implies
+\begin{align}
+  \D\p{G} \O
+  &= \sum_{X\in\set{\z,+,-}} \abs{G_X}^2 \D\p{S_X} \O
+  + \sum_\mu \p{G_\bmu^* G_\mu S_\mu \O S_\mu
+    + G_\z^* G_\mu S_\z \O S_\mu
+    + G_\bmu^* G_\z S_\mu \O S_\z}
+  \notag \\
+  &\qquad -\f12 \sum_\mu\p{G_\bmu^* G_\mu \sp{S_\mu^2, \O}_+
+    + G_\z^*G_\mu \sp{S_\z S_\mu, \O}_+
+    + G_\bmu^* G_\z \sp{S_\mu S_\z, \O}_+}.
+\end{align}
+In order to compute the effect of this decoherence on general
+collective spin operators, we expand the anti-commutators
+\begin{align}
+  \sp{S_\mu^2, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
+  &= S_\mu^{\ell+2} \sp{\p{2+\mu S_\z}^m+\p{\mu S_\z}^m} S_\bmu^n
+  - 2n S_\mu^{\ell+1} \p{n+2\mu S_\z} \p{1+\mu S_\z}^m S_\bmu^{n-1}
+  \notag \\
+  &\qquad + n\p{n-1} S_\mu^\ell \sp{\p{n-1}\p{n-2}
+    + 2\p{2n-3}\mu S_\z + 4\p{\mu S_\z}^2} \p{\mu S_\z}^m S_\bmu^{n-2},
+  \allowdisplaybreaks \\
+  \sp{S_\z S_\mu, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
+  &= \mu S_\mu^{\ell+1} \sp{\p{\ell+1+\mu S_\z}\p{\mu S_\z}^m
+    + \p{n+1+\mu S_\z} \p{1+\mu S_\z}^m } S_\bmu^n \notag \\
+  &\qquad - \mu n S_\mu^\ell \sp{n \p{n-1}
+    + \p{3n-1}\mu S_\z + 2\p{\mu S_\z}^2} \p{\mu S_\z}^m S_\bmu^{n-1},
+  \allowdisplaybreaks \\
+  \sp{S_\mu S_\z, S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_+
+  &= \mu S_\mu^{\ell+1} \sp{\p{\ell+\mu S_\z}\p{\mu S_\z}^m
+    + \p{n+\mu S_\z} \p{1+\mu S_\z}^m} S_\bmu^n \notag \\
+  &\qquad - \mu n S_\mu^\ell \sp{\p{n-1}^2
+    + 3\p{n-1}\mu S_\z + 2\p{\mu S_\z}^2} \p{\mu S_\z}^m S_\bmu^{n-1}.
+\end{align}
+Collecting terms and defining
+\begin{align}
+  G_{\z,\mu}^{(\pm)}
+  &\equiv \f12\p{G_\z^* G_\mu \pm G_\bmu^* G_\z},
+  \allowdisplaybreaks \\
+  \tilde L_{\ell mn;\mu}^{(G)}
+  &\equiv \mu \sp{\p{\ell-n+\f12} G_{\z,\mu}^{(+)}
+    + \p{\ell+\f12} G_{\z,\mu}^{(-)}}
+  S_\mu^{\ell+1} \p{1+\mu S_\z}^m S_\bmu^n \notag \\
+  &\qquad -\mu \sp{\p{\ell-n+\f12} G_{\z,\mu}^{(+)}
+    + \p{n+\f12} G_{\z,\mu}^{(-)}}
+  S_\mu^{\ell+1} \p{\mu S_\z}^m S_\bmu^n \notag \\
+  &\qquad + \mu G_{\z,\mu}^{(-)}
+  S_\mu^{\ell+1} \mu S_\z
+  \sp{\p{1+\mu S_\z}^m - \p{\mu S_\z}^m} S_\bmu^n,
+  \allowdisplaybreaks \\
+  \tilde M_{\ell mn;\mu}^{(G)}
+  &= -\mu n\p{n-1} \sp{\p{\ell-n+\f12} G_{\z,\mu}^{(+)}
+    + \p{\ell-\f12} G_{\z,\mu}^{(-)}}
+  S_\mu^\ell \p{\mu S_\z}^m S_\bmu^{n-1} \notag \\
+  &\qquad - 2\mu n \sp{\p{\ell-n+\f12} G_{\z,\mu}^{(+)}
+    + \p{\ell+\f12n-1} G_{\z,\mu}^{(-)}}
+  S_\mu^\ell \p{\mu S_\z}^{m+1} S_\bmu^{n-1} \notag \\
+  &\qquad - 2\mu n G_{\z,\mu}^{(-)}
+  S_\mu^\ell \p{\mu S_\z}^{m+2} S_\bmu^{n-1},
+  \allowdisplaybreaks \\
+  \tilde P_{\ell mn;\mu}
+  &\equiv -\f12 S_\mu^{\ell+2}
+  \sp{\p{2+\mu S_\z}^m - 2\p{1+\mu S_\z}^m + \p{\mu S_\z}^m}
+  S_\bmu^n \notag \\
+  &\qquad + n S_\mu^{\ell+1} \sp{\p{n+2\mu S_\z} \p{1+\mu S_\z}^m
+    - \p{n-1+2\mu S_\z} \p{\mu S_\z}^m}
+  S_\bmu^{n-1} \notag \\
+  &\qquad -n\p{n-1} S_\mu^\ell
+  \sp{\f12\p{n-1}\p{n-2} + \p{2n-3}\mu S_\z + 2\p{\mu S_\z}^2}
+  \p{\mu S_\z}^m S_\bmu^{n-2},
+  \allowdisplaybreaks \\
+  \tilde Q_{\ell mn;\mu}^{(G)}
+  &\equiv G_\bmu^* G_\mu \tilde P_{\ell mn;\mu}
+  + \tilde L_{\ell mn;\mu}^{(G)}
+  + \tilde M_{\ell mn;\mu}^{(G)},
+\end{align}
+we then have
+\begin{align}
+  \D\p{G} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
+  = \sum_{X\in\set{\z,+,-}} \abs{G_X}^2
+  \D\p{S_X} \p{S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}
+  + \tilde Q_{\ell mn;\mu}^{(G)}
+  + \sp{\tilde Q_{nm\ell;\mu}^{(G)}}^\dag.
+\end{align}
+Note that the sum $\ell+m+n$ for operators
+$S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$ grows by one if $G_\mu\ne0$ or
+$G_\bmu\ne0$, and does not grow otherwise.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Initial conditions}
+\label{sec:initial_conditions}
+
+Here we compute the expectation values of collective spin operators
+with respect to spin-polarized (also Gaussian, or spin-coherent)
+states.  These states are parameterized by polar and azimuthal angles
+$\theta\in[0,\pi)$, $\phi\in[0,2\pi)$, and lie within the Dicke
+manifold spanned by states
+$\ket{k}\propto S_+^{S+k}\ket{\dn}^{\otimes N}$ with $S\equiv N/2$ and
+$S_\z\ket{k}=k\ket{k}$:
+\begin{align}
+  \ket{\theta,\phi}
+  \equiv \sp{\cos\p{\theta/2} e^{-i\phi/2} \ket\up
+    + \sin\p{\theta/2} e^{i\phi/2} \ket\dn}^{\otimes N}
+  = \sum_{k=-S}^S { N \choose S+k }^{1/2}
+  \cos\p{\theta/2}^{S+k} \sin\p{\theta/2}^{S-k} e^{-ik\phi} \ket{k}.
+\end{align}
+We can likewise expand, within the Dicke manifold,
+\begin{align}
+  S_\z = \sum_{k=-S}^S k \op{k},
+  &&
+  S_\mu = \sum_{k=-S+\delta_{\mu,-1}}^{S-\delta_{\mu,1}}
+  g_\mu\p{k} \op{k+\mu}{k}
+  = \sum_{k=-S+\delta_{\bmu,-1}}^{S-\delta_{\bmu,1}}
+  g_\bmu\p{k} \op{k}{k+\bmu},
+\end{align}
+where $\bmu\equiv-\mu\in\set{+1,-1}$ and
+\begin{align}
+  g_\mu\p{k} \equiv \sqrt{\p{S-\mu k}\p{S+\mu k+1}},
+\end{align}
+which implies
+\begin{align}
+  S_\mu^\ell \p{\mu S_\z^m} S_\bmu^n
+  &= \sum_{k=-S+\delta_{\mu,-1}\max\set{\ell,n}}
+  ^{S-\delta_{\mu,1}\max\set{\ell,n}} \p{\mu k}^m
+  \sp{\prod_{p=0}^{\ell-1} g_\mu\p{k+\mu p}}
+  \sp{\prod_{q=0}^{n-1} g_\mu\p{k+\mu q}}
+  \op{k+\mu\ell}{k+\mu n} \\
+  &= \sum_{\mu k=-\mu S-\delta_{\mu,-1}\max\set{\ell,n}}
+  ^{\mu S-\delta_{\mu,1}\max\set{\ell,n}} \p{\mu k}^m
+  \f{\p{S-\mu k}!}{\p{S+\mu k}!}
+  \sp{\f{\p{S+\mu k+\ell}!}{\p{S-\mu k-\ell}!}
+    \f{\p{S+\mu k+n}!}{\p{S-\mu k-n}!}}^{1/2}
+  \op{k+\mu\ell}{k+\mu n} \\
+  &= \sum_{k=-S}^{S-\max\set{\ell,n}} k^m
+  \f{\p{S-k}!}{\p{S+k}!}
+  \sp{\f{\p{S+k+\ell}!}{\p{S-k-\ell}!}
+    \f{\p{S+k+n}!}{\p{S-k-n}!}}^{1/2}
+  \op{\mu\p{k+\ell}}{\mu\p{k+n}}.
+\end{align}
+This expansion allows us to compute the expectation value
+\begin{align}
+  \bk{\theta,\phi|S_\mu^\ell \p{\mu S_\z^m} S_\bmu^n|\theta,\phi}
+  &= e^{i\phi \mu\p{\ell-n}} N! \sum_{k=-S}^{S-\max\set{\ell,n}}
+  \f{k^m \p{S-k}! f_{\mu\ell n}\p{k,\theta}}
+  {\p{S+k}! \p{S-k-\ell}! \p{S-k-n}!} \\
+  &= e^{i\phi \mu\p{\ell-n}} \p{-1}^m N! \sum_{k=0}^{N-\max\set{\ell,n}}
+  \f{\p{S-k}^m \p{N-k}! \tilde f_{\mu\ell n}\p{k,\theta}}
+  {k! \p{N-k-\ell}! \p{N-k-n}!}
+\end{align}
+where
+\begin{align}
+  f_{\mu\ell n}\p{k,\theta}
+  \equiv \cos\p{\theta/2}^{N+\mu\p{2k+\ell+n}}
+  \sin\p{\theta/2}^{N-\mu\p{2k+\ell+n}},
+\end{align}
+\begin{align}
+  \tilde f_{\mu\ell n}\p{k,\theta}
+  \equiv f_{\mu\ell n}\p{k-S,\theta}
+  = \cos\p{\theta/2}^{2N\delta_{\mu,-1}+\mu\p{2k+\ell+n}}
+  \sin\p{\theta/2}^{2N\delta_{\mu,1}-\mu\p{2k+\ell+n}}.
+\end{align}
+Defining the states
+\begin{align}
+  \ket{+\Z} \equiv \ket{0,0} = \ket\up^{\otimes N}, &&
+  \ket{-\Z} \equiv \ket{\pi,0} = \ket\dn^{\otimes N}, &&
+  \ket\X \equiv \ket{\pi/2,0}
+  = \p{\f{\ket\up+\ket\dn}{\sqrt2}}^{\otimes N},
+\end{align}
+some particular expectation values of interest are
+\begin{align}
+  \bk{\nu\Z|S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n|\nu\Z}
+  = \delta_{\ell n} \times
+  \begin{cases}
+    \p{S-n}^m \f{N! n!}{\p{N-n}!} & \mu = \nu, \\
+    \delta_{n,0} \p{-S}^m & \mu \ne \nu,
+  \end{cases},
+\end{align}
+and
+\begin{align}
+  \bk{\X|S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n|\X}
+  = \p{-1}^m \f{N!}{2^N} \sum_{k=0}^{N-\max\set{\ell,n}}
+  \f{\p{S-k}^m \p{N-k}!}{k!\p{N-k-\ell}!\p{N-k-n}!}.
+\end{align}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Computing correlators with the truncated short-time (TST)
+  expansion}
+\label{sec:tutorial}
+
+Here we provide a pedagogical tutorial for computing correlators using
+the truncated short-time TST expansion.  For concreteness, we
+nominally consider $N$ spins evolving under the one-axis twisting
+(OAT) Hamiltonian
+\begin{align}
+  H_{\t{OAT}} = \chi S_\z^2,
+\end{align}
+additionally subject to spontaneous single-spin decay at rate
+$\gamma_-$, with jump operators $\J_-=\set{s_-^{(j)}:j=1,2,\cdots,N}$.
+The equation of motion for a Heisenberg operator
+$\p{S_+^\ell S_\z^m S_-^n}\p{t}$ is
+\begin{align}
+  \f{d}{dt} \Bk{S_+^\ell S_\z^m S_-^n}
+  = i\chi\Bk{\sp{S_\z^2, S_+^\ell S_\z^m S_-^n}_-}
+  + \gamma_- \Bk{\D\p{\J_-}\p{S_+^\ell S_\z^m S_-^n}},
+  \label{eq:tutorial_EOM}
+\end{align}
+where we have suppressed the explicit time dependence of operators for
+brevity.  Using the results in appendices \ref{sec:general_product}
+and \ref{sec:decay_single} respectively to evaluate the commutator
+$\sp{S_\z^2, S_+^\ell S_\z^m S_-^n}_-$ and dissipator
+$\D\p{\J_-} \p{S_+^\ell S_\z^m S_-^n}$ in \eqref{eq:tutorial_EOM}, we
+can expand
+\begin{multline}
+  \f{d}{dt} \bk{S_+^\ell S_\z^m S_-^n} \\
+  = i\chi \bk{\p{\ell-n} S_+^\ell \p{\ell+n+2S_\z} S_\z^m S_-^n}
+  + \gamma_- \Bk{S_+^\ell \sp{\p{S+S_\z}\p{-1+S_\z}^m
+      - \p{S+\f{\ell+n}{2}+S_\z} S_\z^m} S_-^n}.
+  \label{eq:tutorial_EOM_expanded}
+\end{multline}
+In practice, we do not want to keep track of such an expansion by
+hand, especially in the case of e.g.~the two-axis twisting (TAT) and
+twist-and-turn (TNT) models with more general types of decoherence,
+for which the analogue of \eqref{eq:tutorial_EOM_expanded} may take
+several lines just to write out in full.  Defining the operators
+$\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$ with
+$\v m\equiv\p{m_+,m_\z,m_-}$ for shorthand, we note that the vector
+space spanned by $\set{\S_{\v m}}$ is closed under time evolution.  We
+therefore expand
+\begin{align}
+  \f{d}{dt} \bk{\S_{\v n}}
+  = \bk{T \S_{\v n}}
+  = \sum_{\v m} \bk{\S_{\v m}} T_{\v m\v n},
+  \label{eq:tutorial_EOM_general}
+\end{align}
+where $T$ is a superoperator that generates time evolution for
+Heisenberg operators.  In the present example, the matrix elements
+$T_{\v m\v n}\in\C$ of $T$ are defined by
+\eqref{eq:tutorial_EOM_expanded} and \eqref{eq:tutorial_EOM_general}.
+For any Hamiltonian $H$ with decoherence characterized by sets of jump
+operators $\J$ and decoherence rates $\gamma_\J$, the matrix elements
+$T_{\v m\v n}$ are more generally defined by
+\begin{align}
+  T \S_{\v n}
+  = i\sp{H,\S_{\v n}}_-
+  + \sum_\J \gamma_\J \D\p{\J} \S_{\v n}
+  = \sum_{\v m} \S_{\v m} T_{\v m\v n}.
+\end{align}
+The results in Appendices \ref{sec:general_product},
+\ref{sec:decoherence_single}, and \ref{sec:decoherence_collective} can
+be used to write model-agnostic codes that compute matrix elements
+$T_{\v m\v n}$, taking a particular Hamiltonian $H$ and decoherence
+processes $\set{\p{\J,\gamma_\J}}$ as inputs.
+
+In order to compute a quantity such as spin squeezing, we need to
+compute correlators of the form $\bk{\S_{\v n}\p{t}}$, where for
+clarity we will re-introduce the explicit time dependence of
+Heisenberg operators $\S_{\v n}\p{t}$.  The order-$M$ truncated
+short-time (TST) expansion takes
+\begin{align}
+  \bk{\S_{\v n}\p{t}}
+  = \bk{e^{tT}\S_{\v n}\p{0}}
+  = \sum_{k\ge0} \f{t^k}{k!} \bk{T^k\S_{\v n}\p{0}}
+  = \sum_{k\ge0} \f{t^k}{k!}
+  \sum_{\v m} \bk{\S_{\v m}\p{0}} T^k_{\v m\v n}
+  \to \sum_{k=0}^M \f{t^k}{k!}
+  \sum_{\v m} \bk{\S_{\v m}\p{0}} T^k_{\v m\v n},
+  \label{eq:tutorial_TST}
+\end{align}
+where $T^k_{\v m\v n}$ are matrix elements of the $k$-th time
+derivative operator $T^k$, given by
+\begin{align}
+  T^0_{\v m\v n} \equiv
+  \begin{cases}
+    1 & \v m = \v n, \\
+    0 & \t{otherwise}
+  \end{cases},
+  &&
+  T^1_{\v m\v n} \equiv T_{\v m\v n},
+  &&
+  T^{k>1}_{\v m\v n}
+  \equiv \sum_{\v p_1,\v p_2,\cdots,\v p_{k-1}}
+  T_{\v m\v p_{k-1}} \cdots T_{\v p_3\v p_2}
+  T_{\v p_2\v p_1} T_{\v p_1\v n}.
+\end{align}
+Matrix elements $T^k_{\v m\v n}$ and initial-time expectation values
+$\bk{\S_{\v m}\p{0}}$ are thus computed as needed for any particular
+correlator $\bk{\S_{\v n}\p{t}}$ of interest, and combined according
+to \eqref{eq:tutorial_TST}.  Note that initial-time expectation values
+$\bk{\S_{\v m}\p{0}}$ are an {\it input} to the TST expansion, and
+need to be computed separately for any initial state of interest;
+expectation values with respect to spin-polarized (Gaussian) states
+are provided in Appendix \ref{sec:initial_conditions}.  In practice,
+we further collect terms in \eqref{eq:tutorial_TST} to write
+\begin{align}
+  \bk{\S_{\v n}\p{t}} \to \sum_{k=0}^M c_{\v n k} t^k,
+  &&
+  c_{\v n k}
+  \equiv \f1{k!} \sum_{\v m} \bk{\S_{\v m}\p{0}} T^k_{\v m\v n},
+  \label{eq:tutorial_corr}
+\end{align}
+where $c_{\v n k}$ are time-independent coefficients for the expansion
+of $\bk{\S_{\v n}\p{t}}$.  After computing the coefficients
+$c_{\v n k}$, there is only negligible computational overhead to
+compute the correlator $\bk{\S_{\v n}\p{t}}$ for any time $t$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Analytical results for the one-axis twisting model}
+\label{sec:OAT}
+
+The one-axis twisting (OAT) Hamiltonian for $N$ spin-1/2 particles
+takes the form
+\begin{align}
+  H_{\t{OAT}}
+  = \chi S_\z^2
+  = \f12 \chi \sum_{j<k} \sigma_\z^{(j)} \sigma_\z^{(k)} + \f14 N \chi,
+\end{align}
+where $\sigma_\z^{(j)}$ represents a Pauli-$z$ operator acting on spin
+$j$.  This model is a special case of the zero-field Ising Hamiltonian
+previously solved in Ref.~[\citenum{foss-feig2013nonequilibrium}] via
+exact, analytical treatment of the quantum trajectory Monte Carlo
+method for computing expectation values.  The solution therein
+accounts for coherent evolution in addition to decoherence via
+uncorrelated single-spin decay, excitation, and dephasing respectively
+at rates $\gamma_-$, $\gamma_+$, and $\gamma_\z$ (denoted by
+$\Gamma_{\t{ud}}$, $\Gamma_{\t{du}}$, and $\Gamma_{\t{el}}$ in
+Ref.~[\citenum{foss-feig2013nonequilibrium}]).  Letting $S\equiv N/2$
+and $\mu,\nu\in\set{+1,-1}$, we adapt expectation values computed in
+Ref.~[\citenum{foss-feig2013nonequilibrium}] for the initial state
+$\ket\X\propto\p{\ket\up+\ket\dn}^{\otimes N}$ with
+$S_\x\ket\X=S\ket\X$ evolving under $H_{\t{OAT}}$, finding
+\begin{align}
+  \bk{S_+\p{t}}
+  &= S e^{-\kappa t} \Phi\p{\chi,t}^{N-1},
+  \label{eq:S+_OAT} \\
+  \bk{\p{S_\mu S_\z}\p{t}}
+  &= -\f{\mu}{2}\bk{S_\mu\p{t}}
+  + S \p{S-\f12} e^{-\kappa t}
+  \Psi\p{\mu\chi,t} \Phi\p{\chi,t}^{N-2}, \\
+  \bk{\p{S_\mu S_\nu}\p{t}}
+  &= \p{1-\delta_{\mu\nu}} \p{S + \mu\bk{S_\z\p{t}}}
+  + S \p{S-\f12} e^{-2\kappa t} \Phi\p{\sp{\mu+\nu}\chi,t}^{N-2},
+  \label{eq:SS+-_OAT}
+\end{align}
+where
+\begin{align}
+  \Phi\p{X,t}
+  \equiv e^{-\lambda t}
+  \sp{\cos\p{\omega_X t} + \f{\lambda}{\omega_X} \sin\p{\omega_X t}},
+  &&
+  \Psi\p{X,t}
+  \equiv e^{-\lambda t} \p{\f{\Delta+iX}{\omega_X}} \sin\p{\omega_X t},
+\end{align}
+for
+\begin{align}
+  \kappa \equiv \f12 \p{\gamma_+ + \gamma_- + \gamma_\z},
+  &&
+  \lambda \equiv \f12 \p{\gamma_+ + \gamma_-},
+  &&
+  \Delta \equiv \gamma_+ - \gamma_-,
+  &&
+  \omega_X \equiv \sqrt{X^2-\lambda^2-iX\Delta}.
+\end{align}
+In order to compute spin squeezing as measured by the Ramsey squeezing
+parameter $\xi^2$ defined in \eqref{eq:squeezing}, we additionally
+need analytical expressions for $\bk{S_\z\p{t}}$ and
+$\bk{S_\z^2\p{t}}$.  As these operators commute with both the OAT
+Hamiltonian and the single-spin operators $\sigma_\z^{(j)}$, their
+evolution is governed entirely by decay-type decoherence (see Appendix
+\ref{sec:decay_single}), which means
+\begin{align}
+  \f{d}{dt} \bk{S_\z\p{t}}
+  &= S\p{\gamma_+-\gamma_-} - \p{\gamma_++\gamma_-} \bk{S_\z\p{t}},
+  \\
+  \f{d}{dt} \bk{S_\z^2\p{t}}
+  &= S\p{\gamma_++\gamma_-}
+  + 2\p{S-\f12}\p{\gamma_+-\gamma_-} \bk{S_\z\p{t}}
+  - 2 \p{\gamma_++\gamma_-} \bk{S_\z^2\p{t}}.
+\end{align}
+The initial conditions $\bk{S_\z\p{0}}=0$ and $\bk{S_\z^2\p{0}}=S/2$
+then imply
+\begin{align}
+  \bk{S_\z\p{t}} = S\p{\f{\gamma_+-\gamma_-}{\gamma_++\gamma_-}}
+  \p{1-e^{-\p{\gamma_+ + \gamma_-} t}},
+  &&
+  \bk{S_\z^2\p{t}}
+  = \f12 S + S \p{S-\f12} \p{\f{\bk{S_\z\p{t}}}{S}}^2.
+  \label{eq:Sz_OAT}
+\end{align}
+With appropriate assumptions about the relevant sources of
+decoherence, the expectation values in
+\eqref{eq:S+_OAT}--\eqref{eq:SS+-_OAT} and \eqref{eq:Sz_OAT} are
+sufficient to compute the spin squeezing parameter $\xi^2$ in
+\eqref{eq:squeezing} at any time throughout evolution of the initial
+state $\ket\X$ under $H_{\t{OAT}}$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Heisenberg operators in open quantum systems}
+\label{sec:noise}
+
+Here we explain the origin and character of the mean-zero ``noise''
+operators $\E_\O\p{t}$ that appear in the expansion of a Heisenberg
+operator $\O\p{t}=\sum_{\v m}\O_{\v m}\p{t}\S_{\v m}+\E_\O\p{t}$ with
+time-dependent coefficients $\O_{\v m}\p{t}$ for time-independent
+Schr\"odinger operators $\hat\S_{\v m}$.  Our discussion should
+clarify why noise operators play no role in our calculation of
+expectation values of the form $\bk{\O\p{t}}$ and $\bk{\O\Q\p{t}}$,
+despite the fact that noise operators generally {\it do} need to be
+considered in the calculation of more general multi-time correlators
+in open quantum systems\cite{blocher2019quantum}.
+
+In any closed quantum system with initial state $\rho$ and propagator
+$U\p{t}$, such that the state at time $t$ is
+$\rho\p{t}\equiv U\p{t}\rho U^\dag\p{t}$, time-dependent Heisenberg
+operators $\O\p{t}$ are uniquely defined from time-independent
+Schr\"odinger operators $\O$ by
+\begin{align}
+  \bk{\O\p{t}} \equiv \tr\sp{\rho\p{t}\O} = \tr\sp{\rho\O\p{t}}.
+  \label{eq:heisenberg_condition}
+\end{align}
+Enforcing \eqref{eq:heisenberg_condition} for {\it arbitrary} initial
+states $\rho$ forces $\O\p{t}=U^\dag\p{t}\O U\p{t}$.  In an open
+quantum system, however, the definition of a Heisenberg operator is
+not so straightforward.  Open systems can often be understood as
+subsystems of a larger closed system.  Consider therefore an open
+system $S$ with environment $E$, a joint initial state $\rho_{SE}$,
+and propagator $U_{SE}\p{t}$.  The reduced state $\rho_S\p{t}$ of $S$
+at time $t$ is
+\begin{align}
+  \rho_S\p{t}
+  \equiv \tr_E\sp{\rho_{SE}\p{t}}
+  = \tr_E\sp{U_{SE}\p{t} \rho_{SE} U_{SE}^\dag\p{t}}
+  \equiv \U_{\ul S}\p{t} \rho_S,
+  \label{eq:state_S}
+\end{align}
+where $\rho_S\equiv\rho_S\p{0}$ is a time-independent state of $S$ in
+the Heisenberg picture, $\ul S$ denotes the space of operators on $S$,
+and the quantum channel $\U_{\ul S}\p{t}$ has the
+decomposition\cite{rivas2012time}
+\begin{align}
+  \U_{\ul S}\p{t} \rho_S
+  = \sum_j \U_S^{(j)}\p{t} \rho_S {\U_S^{(j)}}^\dag\p{t}
+\end{align}
+with ordinary operators $\U_S^{(j)}\p{t}$ on $S$.  We can therefore
+expand
+\begin{align}
+  \bk{\O_S\p{t}}
+  = \tr\sp{\rho_S\p{t}\O_S}
+  = \tr\sp{\U_{\ul S}\p{t} \rho_S \O_S}
+  = \tr\sp{\rho_S \U_{\ul S}^\dag\p{t} \O_S}
+  = \tr\sp{\rho_S \ul{\O_S}\p{t}}
+  = \bk{\ul{\O_S}\p{t}},
+  \label{eq:heisenberg_correlator}
+\end{align}
+where $\U_{\ul S}^\dag\p{t}$ is the adjoint map of $\U_{\ul S}\p{t}$
+(with respect to a trace inner product between operators on $S$), and
+we define the time-dependent operator
+\begin{align}
+  \ul{\O_S}\p{t}
+  \equiv \U_{\ul S}^\dag\p{t} \O_S
+  = \sum_j {\U_S^{(j)}}^\dag\p{t} \O_S \U_S^{(j)}\p{t}.
+  \label{eq:heisenberg_wrong}
+\end{align}
+We thus find that substituting $\ul{\O_S}\p{t}$ in place of
+$\O_S\p{t}$ suffices for the calculation of correlators
+$\bk{\O_S\p{t}}$, thereby accounting for the validity of the equation
+of motion in \eqref{eq:EOM}.  As we show below, this substitution also
+suffices for the calculation of two-time correlators of the form
+$\bk{\O_S\Q_S\p{t}}$ when the environment $E$ is Markovian.
+
+The problem with {\it defining} Heisenberg operators $\O_S\p{t}$ by
+$\ul{\O_S}\p{t}$ only becomes evident when considering products of
+Heisenberg operators.  One would like for the product of two
+Heisenberg operators $\O_S\p{t}$ and $\Q_S\p{t}$ to satisfy
+$\O_S\p{t}\Q_S\p{t}=\p{\O_S\Q_S}\p{t}$.  This intuition can be
+formalized by observing that
+\begin{align}
+  \bk{\O_S\p{t}}
+  = \tr\sp{\rho_{SE}\p{t} \p{\O_S\otimes\1_E}}
+  = \tr\sp{\rho_{SE} \p{\O_S\otimes\1_E}\p{t}}
+  = \bk{\p{\O_S\otimes\1_E}\p{t}},
+  \label{eq:heisenberg_expansion}
+\end{align}
+where $\1_E$ is the identity operator on $E$, expectation values of
+Heisenberg operators on system $A\in\set{S,E,SE}$ are taken with
+respect to the state $\rho_A$, and
+\begin{align}
+  \p{\O_S\otimes\1_E}\p{t}
+  \equiv U_{SE}^\dag\p{t} \p{\O_S\otimes\1_E} U_{SE}\p{t}.
+  \label{eq:heisenberg_extension}
+\end{align}
+By expanding Heisenberg operators similarly to
+\eqref{eq:heisenberg_expansion} and \eqref{eq:heisenberg_extension},
+we then find
+\begin{align}
+  \bk{\O_S\p{t}\Q_S\p{t}}
+  = \bk{\p{\O_S\otimes\1_E}\p{t} \p{\Q_S\otimes\1_E}\p{t}}
+  = \bk{\p{\O_S\Q_S\otimes\1_E}\p{t}}
+  = \bk{\p{\O_S\Q_S}\p{t}}.
+\end{align}
+The expression in \eqref{eq:heisenberg_wrong}, however, makes it clear
+that generally
+$\ul{\O_S}\p{t} \ul{\Q_S}\p{t}\ne\p{\ul{\O_S\Q_S}}\p{t}$.  To correct
+for this discrepancy, we define
+\begin{align}
+  \O_S\p{t} \equiv \ul{\O_S}\p{t} + \E_{\O_S}\p{t}
+  \label{eq:noise_def}
+\end{align}
+in terms of new ``noise'' operators $\E_{\O_S}\p{t}$ that are
+essentially defined to enforce the consistency of operator products
+such as $\O_S\p{t}\Q_S\p{t}=\p{\O_S\Q_S}\p{t}$. Self-consistency
+forces noise operators to be mean-zero, as
+\begin{align}
+  \bk{\E_{\O_S}\p{t}} = \bk{\O_S\p{t}} - \bk{\ul{\O_S}\p{t}} = 0.
+\end{align}
+Furthermore, if the environment $E$ is Markovian, then noise operators
+are also uncorrelated with initial-time observables,
+i.e.~$\bk{\O_S\E_{\Q_S}\p{t}}=0$, which means that noise operators can
+be neglected in the calculation of two-time correlators of the form
+$\bk{\O_S\Q_S\p{t}}$.  To see why, we observe that a Markovian
+environment is essentially defined to satisfy
+\begin{align}
+  \rho_{SE}\p{t} = U_{SE}\p{t} \rho_{SE} U_{SE}^\dag\p{t}
+  \approx \rho_S\p{t}\otimes\rho_E
+  = \U_{\ul S}\p{t} \rho_S\otimes\rho_E,
+  \label{eq:markov}
+\end{align}
+with $\rho_E$ a time-independent steady state of the environment.  If
+we enforce \eqref{eq:markov} for all states $\rho_S$, e.g.~the
+maximally mixed state $\rho_S^{(1)}\propto\1_S$ and
+$\rho_S^{(2)}\equiv\rho_S^{(1)}+\O_S$ with $\O_S$ any traceless
+operator on $S$ with squared norm
+$\norm{\O_S}^2\equiv\tr\p{\O_S^\dag\O_S}\le\tr\1_S$, then by linearity
+we find that
+\begin{align}
+  U_{SE}\p{t} \p{\1_S\otimes\rho_E} U_{SE}^\dag\p{t}
+  \approx \U_{\ul S}\p{t} \1_S\otimes\rho_E,
+  &&
+  U_{SE}\p{t} \p{\O_S\otimes\rho_E} U_{SE}^\dag\p{t}
+  \approx \U_{\ul S}\p{t} \O_S\otimes\rho_E,
+\end{align}
+which implies that the Markov approximation \eqref{eq:markov} holds
+even if we replace $\rho_S$ by any operator on $S$, and in particular
+\begin{align}
+  U_{SE}\p{t} \rho_{SE} \p{\O_S\otimes\1_E} U_{SE}^\dag\p{t}
+  = U_{SE}\p{t} \p{\rho_S\O_S\otimes\rho_E} U_{SE}^\dag\p{t}
+  \approx \U_{\ul S}\p{t}\p{\rho_S\O_S}\otimes\rho_E.
+  \label{eq:makov_op}
+\end{align}
+We can therefore expand
+\begin{align}
+  \bk{\O_S\Q_S\p{t}}
+  &= \tr\sp{\rho_{SE} \p{\O_S\otimes\1_E}
+    U_{SE}^\dag\p{t} \p{\Q_S\otimes\1_E} U_{SE}\p{t}} \\
+  &= \tr\sp{U_{SE}\p{t} \rho_{SE} \p{\O_S\otimes\1_E}
+    U_{SE}^\dag\p{t} \p{\Q_S\otimes\1_E}},
+\end{align}
+and invoke the Markov approximation in \eqref{eq:makov_op} to find
+that
+\begin{align}
+  \bk{\O_S\Q_S\p{t}}
+  \approx \tr\sp{\U_{\ul S}\p{t}\p{\rho_S\O_S} \Q_S}
+  = \tr\sp{\rho_S \O_S \U_{\ul S}^\dag\p{t} \Q_S}
+  = \bk{\O_S\ul{\Q_S}\p{t}},
+\end{align}
+which implies
+\begin{align}
+  \bk{\O_S\E_{\Q_S}\p{t}}
+  = \bk{\O_S\Q_S\p{t}} - \bk{\O_S\ul{\Q_S}\p{t}}
+  \approx 0.
+\end{align}
+Noise operators thus play no role in the calculation of correlators
+such as $C\p{t}$ in \eqref{eq:two_time}.  In contrast, noise operators
+generally {\it do} play a role in the calculation of multi-time
+correlators of the form
+$\bk{\prod_j\O_S^{(j)}\p{t_j}}$\cite{blocher2019quantum}.
+Furthermore, these calculations generally require additional
+assumptions about the environment.  To keep our discussion simple and
+general, we therefore exclude the effects of noise terms in Section
+\ref{sec:multi_time}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Short-time linear response and two-time correlators}
+\label{sec:linear_response}
+
+Here we discuss the appearance of two-time correlation functions in
+the short-time linear response of correlators to perturbations of a
+Hamiltonian.  Consider an initial Hamiltonian $H$ perturbed by an
+operator $V$ with $\norm{V}\ll\norm{H}$, where $\norm{\O}$ denotes the
+operator norm of $\O$, such that the net Hamiltonian is
+$\tilde H=H+V$.  We denote the generator of Heisenberg time evolution
+under the perturbed (unperturbed) Hamiltonian by $\tilde T$ ($T$).
+These generators are related by
+\begin{align}
+  \tilde T = T + i\ul V
+\end{align}
+where $\ul V$ is a superoperator whose action on operators $\O$ is
+defined by
+\begin{align}
+  \ul V \O \equiv \sp{V,\O}_-.
+\end{align}
+Through quadratic order in the time $t$ and linear order in the
+perturbation $\ul V$, we can say that
+\begin{align}
+  e^{t\tilde T}
+  \approx \f12\sp{e^{tT}, e^{it\ul V}}_+
+  \approx e^{tT} + \f12 it \sp{e^{t T}, \ul V}_+.
+\end{align}
+Defining perturbed and unperturbed Heisenberg operators
+$\tilde\O\p{t}$ and $\O\p{t}$ that respectively satisfy
+$\bk{\tilde\O\p{t}}=\bk{e^{t\tilde T}\O}$ and
+$\bk{\O\p{t}}=\bk{e^{tT}\O}$, we thus find that for sufficiently small
+times $t$ and weak perturbations $V$,
+\begin{align}
+  \bk{\tilde\O\p{t} - \O\p{t}}
+  = \Bk{\p{e^{t\tilde T} - e^{tT}} \O}
+  \approx \f12 i t \p{\bk{\sp{V,\O}_-\p{t}}
+    + \bk{\sp{V,\O\p{t}}_-}}.
+  \label{eq:response}
+\end{align}
+Two-time correlators $\bk{V\O\p{t}}$ and $\bk{\O\p{t}V}$, in addition
+to the expectation values $\bk{\p{V\O}\p{t}}$ and
+$\bk{\p{\O V}\p{t}}$, thus determine the short-time linear response of
+correlators $\bk{\O\p{t}}$ to perturbations $V$ of a Hamiltonian.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \section{Changing operator bases, and a general product for algebras
+%   of permutationally-symmetric operators}
+% \label{sec:new_basis}
+
+% For a collective spin system composed of many small spins, the
+% operators $\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$ contain
+% $\abs{\v m}$-body operators (for $\abs{\v m}\equiv\sum_\mu m_\mu$),
+% but also carry lots of ``baggage'' in the form of $k$-body operators
+% with $k<\abs{\v m}$.  To cut down on this overhead, we try to change
+% our basis for collective spin operators from $\S_{\v m}$ to the
+% purely-$\abs{\v m}$-body operators
+% \begin{align}
+%   \tilde\S_{\v m} \equiv \sum_{\p{\v j;\v m}} P_{\v j},
+%   &&
+%   P_{\v j} \equiv \prod_{\mu,a} s_\mu^{\p{j^\mu_a}},
+%   \label{eq:S_P}
+% \end{align}
+% where $\v m \equiv \p{m_+,m_\z,m_-}\in\mathbb{N}_0^3$ with $m_\mu$ the
+% number of $s_\mu$ operators in each term of $\S_{\v m}$; $\v j$ is an
+% ordered list of indices for the spins addressed by $P_{\v j}$; and the
+% sum over $\p{\v j;\v m}$ denotes a sum over all possible $\v j$ for
+% which all indices in $\v j$ are distinct and
+% $\abs{\set{j^\mu_a}} = m_\mu$, such that $j^\mu_a$ indexes the $a$-th
+% spin addressed by an $s_\mu$ operator in $P_{\v j}$.
+
+% In order to make use of the operators $\tilde S_{\v m}$, we will need
+% to compute products of the form $\tilde S_{\v m} \tilde S_{\v n}$.  We
+% therefore expand such a product into terms that have exactly $s$ spins
+% addressed by both operators:
+% \begin{align}
+%   \tilde\S_{\v m} \tilde\S_{\v n}
+%   = \sum_{s\ge0} \sum_{\p{\v j,\v k;\v m,\v n,s}} P_{\v j} P_{\v k},
+%   &&
+%   \sum_{\p{\v j,\v k;\v m,\v n,s}} X \equiv
+%   \sum_{\substack{\p{\v j;\v m},\p{\v k;\v n} \\
+%       \abs{\set{j_\a^\mu}\cap\set{k_\b^\nu}}=s}} X.
+%   \label{eq:SS_PP}
+% \end{align}
+% Collecting terms in which $r_{\mu\nu}$ of the $s_\mu$ operators in
+% $P_{\v j}$ address the same spin as an $s_\nu$ operator in $P_{\v k}$,
+% we have
+% \begin{align}
+%   \tilde\S_{\v m} \tilde\S_{\v n}
+%   = \sum_{s\ge0} \sum_{\p{\v r;\v m,\v n,s}} f_{\v m\v n\v r}
+%   \sum_{\p{\v J,\v K;\v m,\v n,\v r}} P_{\v J} Q_{\v K},
+%   \label{eq:SS_PQ}
+% \end{align}
+% where the sum over $\p{\v r;\v m,\v n,s}$ denotes a sum over all
+% values of $\v r$ with the restrictions
+% \begin{align}
+%   r_{\mu\nu} \ge 0 ~\forall~ \mu,\nu,
+%   &&
+%   \sum_\nu r_{\mu\nu} \le m_\mu,
+%   &&
+%   \sum_\mu r_{\mu\nu} \le n_\nu,
+%   &&
+%   \sum_{\mu,\nu} r_{\mu\nu} = s;
+%   \label{eq:rest_r}
+% \end{align}
+% the factor $f_{\v m\v n\v r}$ counts the number of ways to pair
+% operators in $P_{\v j}$ and $P_{\v k}$ according to $\v r$, or
+% \begin{align}
+%   f_{\v m\v n\v r}
+%   &\equiv \sp{\prod_{\mu,\nu}
+%     { m_\mu - \sum_{\rho<\nu} r_{\mu\rho} \choose r_{\mu\nu} }
+%     { n_\mu - \sum_{\rho<\nu} r_{\rho\mu} \choose r_{\nu\mu} }}
+%   \sp{\prod_{\mu,\nu} r_{\mu\nu}!} \\
+%   &= \sp{\prod_\mu \f{m_\mu!}{\p{m_\mu-\sum_\rho r_{\mu\rho}}!}
+%     \f{n_\mu!}{\p{n_\mu-\sum_\rho r_{\rho\mu}}!}}
+%   \sp{\prod_{\mu,\nu} r_{\mu\nu}!}^{-1};
+% \end{align}
+% $\v J$ and $\v K$ are respectively ordered lists of indices for spins
+% addressed by one and two single-spin operators in a term of the
+% product $P_{\v j} P_{\v k}$ with fixed $\v r$; the sum over
+% $\p{\v J,\v K;\v m,\v n,\v r}$ denotes a sum over all values of
+% $\v J,\v K$ with the restrictions
+% \begin{align}
+%   \abs{\set{J_a^\mu}}
+%   = m_\mu + n_\mu - \sum_\rho\p{r_{\mu\rho}+r_{\rho\mu}},
+%   &&
+%   \abs{\set{K^{\mu\nu}_a}} = r_{\mu\nu},
+%   &&
+%   \set{J_a^\lambda} \cap \set{K_b^{\mu\nu}} = \varnothing,
+% \end{align}
+% with $\varnothing$ denoting the empty set; and finally, similarly to
+% $P_{\v j}$ in \eqref{eq:S_P} we define
+% \begin{align}
+%   Q_{\v K} \equiv \prod_{\mu,\nu,a}
+%   s_\mu^{(K^{\mu\nu}_a)} s_\nu^{(K^{\mu\nu}_a)}.
+% \end{align}
+% As the sum in \eqref{eq:SS_PQ} is invariant under permutation of the
+% indices in $\v K$, we can safely neglect keeping track of these
+% indices and simply write
+% \begin{align}
+%   Q_{\v K}
+%   = \prod_{\mu,\nu} \prod_{a=1}^{r_{\mu\nu}}
+%   s_\mu^{(K^{\mu\nu}_a)} s_\nu^{(K^{\mu\nu}_a)}
+%   \simeq \bigotimes_{\mu,\nu} \bigotimes_{a=1}^{r_{\mu\nu}} s_\mu s_\nu
+%   = \bigotimes_{\mu,\nu} \bigotimes_{a=1}^{r_{\mu\nu}}
+%   \sum_\rho \eta_{\mu\nu\rho} s_\rho,
+%   \label{eq:Q_K_eta}
+% \end{align}
+% where $\simeq$ denotes equality up to a re-indexing of spins and
+% additional tensor factors of the single-spin identity operator
+% (i.e.~$\1$); we have introduced explicit dependence on
+% $r_{\mu\nu}=\abs{\set{K^{\mu\nu}_a}}$ for brevity; and
+% $\eta_{\mu\nu\rho}$ is a structure constant defined by
+% $s_\mu s_\nu=\sum_\rho\eta_{\mu\nu\rho}s_\rho$.  Unlike the sums over
+% $\mu,\nu$ in most of the work above, the sum over $\rho$ here includes
+% an index for the identity operator $s_0\equiv\1$.  Distributing the
+% product of sums into a sum of products gives us
+% \begin{align}
+%   Q_{\v K}
+%   \simeq \sum_{\v\rho} \bigotimes_{\mu,\nu,a}
+%   \eta_{\mu\nu\rho^{\mu\nu}_a} s_{\rho^{\mu\nu}_a}
+%   = \sum_{\v\rho} \sp{\prod_{\mu,\nu,a}\eta_{\mu\nu\rho^{\mu\nu}_a}}
+%   \bigotimes_{\mu,\nu,a} s_{\rho^{\mu\nu}_a},
+%   \label{eq:Q_K_rho}
+% \end{align}
+% where implicitly $\abs{\set{\rho^{\mu\nu}_a}}=r_{\mu\nu}$, as
+% $\p{\mu,\nu,a}$ essentially index a factor in \eqref{eq:Q_K_eta} and
+% $\rho^{\mu\nu}_a$ indexes one of the terms in that factor.  Letting
+% $\tilde\rho^{\mu\nu}_\kappa$ denote the number of elements in the
+% index list
+% $\p{\rho^{\mu\nu}_1,\rho^{\mu\nu}_2,\cdots,\rho^{\mu\nu}_{r_{\mu\nu}}}$
+% that are equal to $\kappa$, we observe that two terms in
+% \eqref{eq:Q_K_rho} with, say, $\v\rho=\v\rho_1$ and $\v\rho=\v\rho_2$
+% are equal up to a permutation of indices if
+% $\tilde{\v\rho}_1=\tilde{\v\rho}_2$.  The degeneracy (under
+% permutation) of terms in \eqref{eq:Q_K_rho}, i.e.~the number of
+% $\v\rho$ that are consistent with $\tilde{\v\rho}$, is
+% \begin{align}
+%   g_{\tilde{\v\rho}}
+%   \equiv \prod_{\mu,\nu,\kappa}
+%   { r_{\mu\nu} - \sum_{\lambda<\kappa} \tilde\rho^{\mu\nu}_\lambda
+%     \choose \tilde\rho^{\mu\nu}_\kappa }
+%   = \sp{\prod_{\mu,\nu} r_{\mu\nu}!}
+%   \sp{\prod_{\mu,\nu,\kappa} \tilde\rho^{\mu\nu}_\kappa!}^{-1}.
+% \end{align}
+% Defining additionally
+% \begin{align}
+%   \v\eta^{\tilde{\v\rho}}
+%   \equiv \prod_{\mu,\nu,\kappa}
+%   \p{\eta_{\mu\nu\kappa}}^{\tilde\rho^{\mu\nu}_\kappa},
+%   \label{eq:eta_rho}
+% \end{align}
+% we can collect like factors and group together equivalent terms in
+% \eqref{eq:Q_K_rho} to write
+% \begin{align}
+%   Q_{\v K}
+%   \simeq \sum_{\tilde{\v\rho}} g_{\tilde{\v\rho}} \v\eta^{\tilde{\v\rho}}
+%   \bigotimes_{\kappa}
+%   s_\kappa^{\otimes\sum_{\mu,\nu}\tilde\rho^{\mu\nu}_\kappa}
+%   \simeq \sum_{\tilde{\v\rho}} g_{\tilde{\v\rho}} \v\eta^{\tilde{\v\rho}}
+%   \1^{\otimes\sum_{\mu,\nu}\tilde\rho^{\mu\nu}_0}
+%   \otimes \bigotimes_{\kappa\ne0}
+%   s_\kappa^{\otimes\sum_{\mu,\nu}\tilde\rho^{\mu\nu}_\kappa},
+%   \label{eq:Q_K_rho_simp}
+% \end{align}
+% where again implicitly
+% $\sum_\kappa\tilde\rho^{\mu\nu}_\kappa=r_{\mu\nu}$ for consistency
+% with \eqref{eq:Q_K_eta} and \eqref{eq:Q_K_rho}, and we have explicitly
+% factored out the identity operators in each term of $Q_{\v K}$.
+
+% We have now simplified $Q_{\v K}$ sufficiently to substitute it back
+% into \eqref{eq:SS_PQ}, which (removing the tilde from $\tilde{\v\rho}$
+% to simplify notation) gives us
+% \begin{align}
+%   \tilde\S_{\v m} \tilde\S_{\v n}
+%   = \sum_{s\ge0} \sum_{\p{\v r;\v m,\v n,s}} f_{\v m\v n\v r}
+%   \sum_{\p{\v\rho;\v r}} g_{\v\rho} \v\eta^{\v\rho}
+%   c_{\v\ell_{\v m\v n\v r\v\rho}\v\rho}
+%   \tilde\S_{\v\ell_{\v m\v n\v r\v\rho}},
+%   \label{eq:SS_simp}
+% \end{align}
+% where the sum over $\p{\v\rho;\v r}$ denotes a sum over all values of
+% $\v\rho$ and with the restrictions
+% \begin{align}
+%   \abs{\set{\rho^{\mu\nu}_\kappa}} = \abs{\set{s_\kappa}},
+%   &&
+%   \sum_{\kappa} \rho^{\mu\nu}_\kappa = r_{\mu\nu};
+% \end{align}
+% the combinatorial factor
+% \begin{align}
+%   c_{\v\ell\v\rho}
+%   \equiv \f{\p{N - \abs{\v\ell}}!}
+%   {\p{N - \abs{\v\ell} - \sum_{\mu,\nu}\rho^{\mu\nu}_0}!}
+% \end{align}
+% accounts for the number of terms that are equivalent up to a
+% permutation of indices involving the identity operators in
+% \eqref{eq:Q_K_rho_simp}; and the number of $s_\mu$ operators in each
+% term of $\tilde\S_{\v\ell_{\v m\v n\v r\v\rho}}$ is
+% \begin{align}
+%   \ell_{\v m\v n\v r\v\rho;\mu}
+%   = m_\mu + n_\mu - \sum_\kappa \p{r_{\mu\kappa}+r_{\kappa\mu}}
+%   + \sum_{\a,\b} \rho^{\a\b}_\mu.
+% \end{align}
+
+% The sums in \eqref{eq:SS_simp} generally involve a large number of
+% terms; in practice, many of these terms will be equal to zero.  We
+% can, however, impose additional restrictions on these sums in such a
+% way as to throw out all terms that are zero, and keep only those that
+% are not.  Remembering that $s$ counts the number of spins that are
+% addressed by both operators $P_{\v j}$ and $P_{\v k}$ in
+% \eqref{eq:SS_PP}, the first restriction we can impose comes from
+% recognizing that for given $\tilde\S_{\v m},\tilde\S_{\v n}$, the
+% overlap $s$ is bounded as
+% \begin{align}
+%   \max\set{0,\abs{\v m}+\abs{\v n}-N}
+%   \le s \le \min\set{N,\abs{\v m}+\abs{\v n}}.
+% \end{align}
+% In addition, the matrix of structure constants $\eta_{\mu\nu\kappa}$
+% may generally contain zeros that we can preemptively avoid with
+% restrictions on $\v r$ and $\v\rho$.  Recalling that $r_{\mu\nu}$
+% counts the number of operators addressed by
+% $s_\mu s_\nu=\sum_\kappa\eta_{\mu\nu\kappa} s_\kappa$, if
+% $\eta_{\mu\nu\kappa}=0$ for all $\kappa$, i.e.~$s_\mu s_\nu=0$, then
+% any term with $r_{\mu\nu}>0$ will vanish.  Our second restriction on
+% the sums in \eqref{eq:SS_simp} is therefore to fix $r_{\mu\nu}=0$ for
+% all $\mu,\nu$ with $s_\mu s_\nu=0$.  Finally, we can avoid the
+% nullifying effect of zeros from the factor
+% $\v\eta^{\v\rho}=\prod_{\mu,\nu,\kappa}
+% \p{\eta_{\mu\nu\kappa}}^{\rho^{\mu\nu}_\kappa}$ by fixing
+% $\rho^{\mu\nu}_\kappa=0$ for all $\mu,\nu,\kappa$ with
+% $\eta_{\mu\nu\kappa}=0$.  These restrictions are sufficient to ensure
+% that all individual terms in \eqref{eq:SS_simp} are nonzero, although
+% they do not rule out the possibility of several terms canceling out.
+
+% TODO: comment on the use of this new basis vs.~the old one.
 

### 2019-09-20 soc_squeezing/squeezing_paper/main.tex
pushing (re)submitted version of squeezing paper

--- 2019-08-28 soc_squeezing/squeezing_paper/main.tex
@@
-\documentclass[aps,prx,superscriptaddress,notitlepage,twocolumn,longbibliography]{revtex4-1}
+\documentclass[aps,prx,superscriptaddress,notitlepage,twocolumn,longbibliography]{revtex4-2}
 
@@
 We demonstrate that even with realistic experimental imperfections, our protocol may generate $\sim10$--$14$ dB of spin squeezing in $\sim1$ second with $\sim10^2$--$10^4$ atoms.
-This capability exemplifies a new paradigm of using driven non-equilibrium systems to overcome current limitations in quantum metrology, allowing OLCs to enter a new regime of enhanced sensing with correlated quantum states.
+This capability allows OLCs to enter a new era of quantum enhanced sensing using correlated quantum states of driven non-equilibrium systems.
 \end{abstract}
@@
 Optical lattice clocks in particular have seen some of the most impressive developments in recent years, reaching record levels of precision ($\sim 3\times 10^{-19}$)\cite{campbell2017fermidegenerate, marti2018imaging} and accuracy ($\sim 1\times 10^{-18}$)\cite{bloom2014optical, mcgrew2018atomic}.
-These advancements required important breakthroughs, including the capability to cool and trap fermionic alkaline-earth atoms in spin-insensitive potentials\cite{takamoto2003spectroscopy, barber2006direct, ye2008quantum}; the development of ultra-coherent lasers\cite{kessler2012sub40mhzlinewidth, cole2013tenfold, matei2017mu} to fully exploit an ultra-narrow clock transition\cite{ludlow2015optical}; the detailed characterization of inter-atomic interactions\cite{scazza2014observation, cappellini2014direct, Zhang2014}; and, more recently, the preparation of a quantum degenerate gas in a three-dimensional (3D) optical lattice\cite{campbell2017fermidegenerate, marti2018imaging, goban2018emergence}.
+These advancements required important breakthroughs, including the capability to cool and trap fermionic alkaline-earth atoms in spin-insensitive potentials\cite{takamoto2003spectroscopy, barber2006direct, ye2008quantum}; the development of ultra-coherent lasers\cite{kessler2012sub40mhzlinewidth, cole2013tenfold, matei2017mu} to fully exploit an ultra-narrow clock transition\cite{ludlow2015optical}; the detailed characterization of inter-atomic interactions\cite{scazza2014observation, cappellini2014direct, zhang2014spectroscopic}; and, more recently, the preparation of a quantum degenerate gas in a three-dimensional (3D) optical lattice\cite{campbell2017fermidegenerate, marti2018imaging, goban2018emergence}.
 Nonetheless, all improvements in sensing capabilities to date have been based on single-particle control of internal atomic degrees of freedom.
@@
 
-In this work, we describe a scheme that can potentially lead to metrological advances in state-of-the-art optical lattice clocks through direct use of quantum entanglement by harnessing the interplay between nominally undesirable collisions and SOC.
-This scheme is made possible in the weak SOC regime by the formation of an interaction-energy gap that suppresses the SOC-induced population transfer from the exchange-symmetric Dicke manifold (spanned by spin-polarized and thus non-interacting states) to the remainder of Hilbert space.
-To generate spin squeezing, our protocol only requires control over (i) the orientation of the clock laser and (ii) the optical lattice depth.
+In this work, we describe a scheme that can lead to metrological advances in state-of-the-art optical lattice clocks through direct use of quantum entanglement by harnessing the interplay between nominally undesirable collisions and SOC.
+This scheme is made possible in the weak SOC regime by the formation of an interaction-energy gap that suppresses the SOC-induced population transfer from the exchange-symmetric Dicke manifold (spanned by spin-polarized, and thus non-interacting states) to the remainder of Hilbert space.
+Interactions thereby prolong inter-particle spin coherence through a spin-locking effect, which additionally transforms the dephasing effect of SOC into a collective spin squeezing process.
+To generate spin squeezing, our protocol only requires the capability to fix (i) the orientation of the clock laser and (ii) the optical lattice depth.
 These controls are straightforward to incorporate into current 3D clock interrogation sequences without sacrificing atom numbers or coherence times.
-Additionally, we show that by applying a modulated drive from the clock laser, one can further prepare states that saturate the Heisenberg limit $\Delta\theta \sim 1/N$ for phase sensitivity\cite{kitagawa1993squeezed, ma2011quantum, degen2017quantum}.
+Additionally, we show that by applying a modulated drive from the clock laser, one can further prepare states that saturate the Heisenberg limit $\Delta\theta\sim1/N$ for phase sensitivity\cite{kitagawa1993squeezed, ma2011quantum, degen2017quantum}.
+This capability mirrors efforts in other settings, such as NV centers\cite{bauch2018ultralong, aiello2013compositepulse} and trapped ions\cite{burd2019quantum}, to enhance quantum metrology through the use of driven non-equilibrium phenomena.
 
-The intriguing and non-trivial observation reported here is that weakly interacting fermions in a lattice with dynamics described by the Hubbard model --- which thus obey fermionic quantum statistics and interact solely via contact collisions --- can feature all-to-all collective spin interactions that naturally generate spin-squeezed states in a manner robust to realistic experimental imperfections.
-In the spirit of quantum simulation, our work illustrates that short-range interacting systems can be used to emulate  dynamical magnetic behaviors found in arrays of long-range interacting systems (such as arrays of trapped ions\cite{Bohnet2016} or cavity QED systems\cite{Monika2010,norcia2018cavitymediated} with interactions mediated by phonons or photons, respectively), thus enhancing their potential as a metrological resource.
-Although surprising, this result is in line with the recent understandings in quantum research that explore how the interplay between well understood simple phenomena can give rise to emergent new  behaviors.
-In this direction, our work emphasizes that the precise control of interactions, which becomes possible in 3D optical lattice systems with quantized motional degrees of freedom in all three directions, is a key for the implementation of the protocols studied here, which may enable new quantum technologies.
+Despite an abundance of proof-of-principle experiments with entangled states\cite{degen2017quantum, pezze2018quantum}, so far only the remarkable example of LIGO\cite{aasi2013enhanced, abbott2016gw150914} has demonstrated a quantum advantage in a state-of-the-art quantum sensing or measurement system.
+The new generation of 3D optical lattice systems have fully quantized motional degrees of freedom\cite{campbell2017fermidegenerate}, allowing for precise control of collisional interactions.
+We demonstrate how these interactions can naturally give rise to metrologically useful correlated many-body fermionic states, opening a path to not only generate entanglement, but also harness it to achieve a quantum advantage in a world-class sensor.
+Such an advance will ultimately deliver gains to real-world applications including timekeeping, navigation, telecommunication, and our understanding of the fundamental laws of nature\cite{safronova2018search}.
 
-\section{From the Fermi-Hubbard model to one-axis and two-axis twisting}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Spin squeezing with the Fermi-Hubbard model}
 \label{sec:theory}
@@
 
-On its own, the collective Heisenberg term ($\sim\hat{\v S}\c\hat{\v S}$) in Eqn.~\eqref{eq:SpinModel} opens an energy gap $fU$, with $f\equiv N/L$ the filling fraction of spatial modes, between the collective Dicke states $\ket{S=N/2, M_S}$ and the remainder of Hilbert space\cite{rey2008manybody, martin2013quantum, norcia2018cavitymediated,Smale2018} with $S<N/2$.
+On its own, the collective Heisenberg term ($\sim\hat{\v S}\c\hat{\v S}$) in Eqn.~\eqref{eq:SpinModel} opens an energy gap $fU$, with $f\equiv N/L$ the filling fraction of spatial modes, between the collective Dicke states $\ket{S=N/2, M_S}$ and the remainder of Hilbert space\cite{rey2008manybody, martin2013quantum, norcia2018cavitymediated, smale2019observation} with $S<N/2$.
 Here $S$ and $M_S$ respectively label the eigenvalues of the collective spin operators $\hat{\v S}\c\hat{\v S}$ and $\hat{S}_\z$, with eigenvalues $S(S+1)$ for non-negative $S\in\set{N/2,N/2-1,\cdots}$ and $M_S\in\set{-S,-S+1,\cdots,S}$.
@@
 
-The validity of the OAT model in Eqn.~\eqref{eq:SpinModel} relies on two key conditions concerning experimental parameter regimes.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Model validity}
+\label{sec:validity}
+
+The validity of the OAT model in Eqn.~\eqref{eq:H_eff} relies on two key conditions concerning experimental parameter regimes.
 First, the on-site interaction energy $U$ should not be much larger in magnitude than the tunneling rate $J$ (clarified below); otherwise, one cannot assume frozen motional degrees of freedom (i.e.~with atoms pinned to fixed quasi-momentum modes) and map the Fermi-Hubbard model to a spin model.
 Second, the SOC-induced fields $B_q\sim J\sin(\phi/2)$ should be considerably smaller in magnitude than the interaction energy gap $fU$, as otherwise one cannot perturbatively transform SOC into OAT.
-These two conditions can be satisfied by appropriate choices of $U/J$ and the SOC angle $\phi$, which are respectively controlled by tuning the lattice depth and changing the angle between the clock laser and shallow lattice axes [see Fig.~\ref{fig:protocol_schematic}(a)].
+These two conditions can be satisfied by appropriate choices of $U/J$ and the SOC angle $\phi$, which are respectively controlled by tuning the lattice depth and changing the angle between the clock laser and the lattice axes [see Fig.~\ref{fig:protocol_schematic}(a)].
 
@@
 Therein, optimal squeezing achievable under unitary dynamics is provided in dB, i.e.~$-10\log_{10}(\xi_{\t{opt}}^2)$, while the time at which this squeezing occurs is provided in units of the nearest-neighbor tunneling time $2\pi/J$.
-At $f=1$ atom per lattice site, i.e.~half filling of all atomic states in the lowest Bloch band, the spin model [Eqn.~\eqref{eq:SpinModel}] agrees almost exactly with the Fermi-Hubbard (FH) model [Eqn.~\eqref{eq:FermiHubbard}] up through $U/J=8$.
+At $f=1$ atom per lattice site, i.e.~half filling of all atomic states in the lowest Bloch band, the spin model [Eqn.~\eqref{eq:SpinModel}] agrees almost exactly with the Fermi-Hubbard (FH) model [Eqn.~\eqref{eq:FermiHubbard}] up through (and exceeding) $U/J=8$.
 The agreement at half filling ($f=1$) is assisted by Pauli blocking of mode-changing collisions.
 Below half filling ($f=5/6$), these two (FH and spin) models show good agreement at $U/J\lesssim2$, while at $U/J\gtrsim2$ mode-changing collisions start to become relevant and invalidate the frozen-mode assumption of the spin model.
-Note that we chose filling $f=5/6$ to demonstrate that our protocol should work even in this highly hole-doped case; in practice, optimized experiments are capable of achieving fillings closer to $f=1$\cite{brown2017spinimbalance}.
+Note that we chose filling $f=5/6$ to demonstrate that our protocol should work, albeit sub-optimally, even in this highly hole-doped case; in practice, optimized experiments are capable of achieving fillings closer to the optimal $f=1$\cite{brown2017spinimbalance}.
 Interestingly, even with mode-changing collisions the Fermi-Hubbard model exhibits comparable amounts of squeezing to the spin model, and achieves this squeezing in less time.
@@
 In realistic implementations, the Gaussian profile of the laser beams always introduces an additional effective harmonic potential that modifies the translational invariance assumed so far.
-We present a detailed discussion of the role of the harmonic trap in Appendix \ref{App:harmonic_trap}, where we demonstrate that the addition of harmonic confinement barely modifies the achievable spin squeezing with currently accessible trapping frequencies.
-We find that the existence of single-particle localized modes in the lattice with harmonic confinement\cite{Rey2005,Pupillo2006} helps to protect spin squeezing and shifts the optimal parameter window to $U/J\gtrsim 2$.
+We present a detailed discussion of the role of the harmonic trap in Appendix \ref{sec:harmonic_trap}, where we demonstrate that the addition of harmonic confinement barely modifies the achievable spin squeezing with currently accessible trapping frequencies.
+We find that the existence of single-particle localized modes in the lattice with harmonic confinement\cite{rey2005ultracold, pupillo2006extended} helps to protect spin squeezing and shifts the optimal parameter window to $U/J\gtrsim 2$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Two-axis twisting}
+\label{sec:TAT}
 
@@
 
-\section{Experimental preparation of a squeezed state and practical considerations}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Experimental implementation and practical considerations}
 
-Here we discuss specific implementations of the above protocols in the state-of-the-art 3D $^{87}$Sr optical lattice clock (OLC).
-This system has demonstrated the capability to load a quantum degenerate gas into a 3D lattice at the ``magic wavelength'' ($\lambda_{\t{lattice}}=2a\approx813$ nm) for which both the ground ($^1S_0,\dn$) and first excited ($^3P_0,\up$) electronic states of the atoms experience the same optical potential\cite{campbell2017fermidegenerate}.
+Thus far, we have largely considered the general preparation of spin-squeezed states with the Fermi-Hubbard model.
+Here, we discuss the specific implementation of the above protocols in the state-of-the-art 3D $^{87}$Sr optical lattice clock (OLC).
+If successful, such an implementation would (to our knowledge) for the first time break through the proof-of-principle stage of spin squeezing efforts, and achieve a genuine metrological enhancement of a world-class quantum sensor.
+
+As required for our protocol, 3D $^{87}$Sr OLC has demonstrated the capability to load a quantum degenerate gas into a 3D lattice at the ``magic wavelength'' ($\lambda_{\t{lattice}}=2a\approx813$ nm) for which both the ground ($^1S_0,\dn$) and first excited ($^3P_0,\up$) electronic states (i.e.~the ``clock states'') of the atoms experience the same optical potential\cite{campbell2017fermidegenerate}.
 Furthermore, the 3D $^{87}$Sr OLC currently operates at sufficiently low temperatures to ensure vanishing population above the lowest Bloch band, such that its dynamics are governed by the Fermi-Hubbard Hamiltonian [Eqn.~\eqref{eq:FermiHubbard}]\cite{esslinger2010fermihubbard}.
@@
 Lattice depths $V_0$ are normalized to the atomic lattice recoil energy $E_{\t{R}}$, and the upper axis on panels ({\bf b}, {\bf c}) marks values of  $U/J$ at fixed lattice depths.
-In general, TAT achieves more squeezing than OAT for any system size, and achieves this squeezing faster for $N\gtrsim500$ atoms.
+In general, TAT achieves more squeezing than OAT for any system size, and achieves optimal squeezing faster for $N\gtrsim400$ atoms, as denoted by a dotted line in panels ({\bf b}, {\bf c}).
 }
@@
 Optimal squeezing times in the presence of decoherence are generally smaller than the corresponding times shown in Fig.~\ref{fig:optima_coherent}, as decoherence typically degrades squeezing before it reaches the decoherence-free maximum.
-Sample plots of squeezing over time for particular choices of lattice size ($\ell$) and depth ($V_0/E_{\t{R}}$) are provided in the Supplementary Material \cite{supplementary_material}.
+The decoherence considered in this work also limits maximally achievable squeezing to $\sim20$ dB less than the decoherence-free maxima shown in Fig.~\ref{fig:optima_coherent}.
+Sample plots of squeezing over time for particular choices of lattice size ($\ell$) and depth ($V_0/E_{\t{R}}$) are provided in the Supplementary Material\cite{supplementary_material}.
 }
@@
 Due to the fast growth of Heisenberg operators in systems with all-to-all interactions, the latter method is not always capable of simulating up to the optimal squeezing time, and thus only provides a lower bound on the maximal squeezing theoretically obtainable via TAT.
+
 The results in Fig.~\ref{fig:optima_decoherence} show that squeezing via OAT saturates with system size around $N\approx10^3$ ($\ell\approx30$), while TAT allows for continued squeezing gains through $N=10^4$ ($\ell=100$).
 Even with decoherence, our protocol may realistically generate $\sim10$--$14$ dB of spin squeezing in $\sim1$ second with $\sim10^2$--$10^4$ atoms in a 2D section of the lattice, which is compatible with the atom numbers and interrogation times of state-of-the-art optical lattice clocks\cite{campbell2017fermidegenerate, marti2018imaging}.
+This amount of spin squeezing exceeds those reported in the ground-state nuclear spin sublevels of a state-of-the-art ${}^{171}$Yb OLC ($\sim6.5$ dB)\cite{braverman2019nearunitary}.
+While the latter protocol might be used to transfer spin squeezing to the electronic clock state, to date there has been no demonstration of spin squeezing in an optical clock transition.
 
-In addition to light scattering, $p$-wave losses from inelastic ${}^3 P_0$ collisions\cite{martin2013quantum, Zhang2014, Bishof2011} can also degrade the maximum achievable spin squeezing, particularly in shallow lattices.
-More details on $p$-wave losses are discussed in Appendix \ref{App:inelastic_collision}, where we show that operating at lattices deeper than $8 E_{\t{R}}$ may be necessary to suppress the impact of inelastic collisions on spin squeezing, at the cost of slightly increasing  light scattering.
+In addition to light scattering, $p$-wave losses from inelastic ${}^3 P_0$ collisions\cite{martin2013quantum, zhang2014spectroscopic, bishof2011inelastic} can also degrade the maximum achievable spin squeezing, which becomes more pronounced for shallower lattices.
+More details on $p$-wave losses are discussed in Appendix \ref{sec:inelastic_collision}, where we show that operating at lattice depths $V_0\gtrsim 7 E_{\t{R}}$ may be necessary to suppress the impact of inelastic collisions on spin squeezing, at the cost of slightly increasing light scattering.
 
@@
 If, for example, the $-9/2$ and $-7/2$ nuclear spin states are used for the preparation of a squeezed state, then the collective-spin entanglement of atoms can be transferred to electronic states at the end of the spin squeezing protocol with a $\sigma^-$ polarized $\pi$ pulse.
-Such a pulse can transfer $\ket{g, I=-7/2}$ to $\ket{e, I=-9/2}$ without affecting $\ket{g, I=-9/2}$.
+Such a pulse can transfer $\ket{g,-7/2}$ to $\ket{e,-9/2}$ without affecting $\ket{g,-9/2}$, where $g$ and $e$ respectively denote the ground and excited (electronic) clock states.
 
-%The only additional detail that needs to be considered for experimental implementation of our protocol is the fact that SOC needs to be suppressed during clock interrogation.
-In a shallow lattice, SOC can be turned off by use of a second laser with SOC angle $\phi=0$, and a simple pulse sequence that transforms our squeezed state into the appropriate gauge (see Appendix \ref{sec:clock_interrogation}).}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Conclusions}
@@
 We have proposed a new protocol to generate spin squeezing in a fermionic 3D optical lattice clock by combining nominally undesirable atomic collisions with spin-orbit coupling.
-To our knowledge, this is the first proposal to use atomic collisions for entanglement generation in state-of-the art atomic clocks, opening a path for the practical improvement of world-leading quantum sensors using correlated many-body fermionic states.
+To our knowledge, this is the first proposal to use quantum correlations in a many-body fermionic system to push state-of-the-art quantum sensors beyond the independent-particle regime, thereby achieving a genuine quantum advantage.
 Such capability could allow for major improvements in clock sensitivity and bandwidth, enhancing not only traditional timekeeping applications such as measurement standards, navigation (GPS), and telecommunications, but also geodesy and gravitational wave detection, precision tests of fundamental physics, and the search for new physics beyond the standard model\cite{safronova2018search}.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Derivation of the effective one-axis-twisting model}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Two-axis twisting, decoherence, and the residual axial field}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Dynamical decoupling in the TAT protocol}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Clock interrogation after squeezing}
@@
 The simplest way to turn off SOC is thus to increase the lattice depth prior to clock interrogation, taking $J\sim B_q\to0$.
-Increasing the lattice depth to turn off SOC is compatible with the current clock interrogation sequence, but is incompatible with ongoing efforts to mitigate light scattering from the lattice beams (currently the primary source of decoherence in the clock; see Appendix \ref{sec:decoherence}) by use of a shallow lattice\cite{hutson2019engineering}.
+Increasing the lattice depth to turn off SOC is compatible with the current clock interrogation sequence, but is incompatible with ongoing efforts to mitigate light scattering from the lattice beams (currently the primary source of decoherence in the clock; see Appendix \ref{sec:decoherence}) by using shallower lattices\cite{hutson2019engineering}.
 We thus devote the rest of this section to discussing strategies for turning off SOC that are compatible with using the same lattice depth for clock interrogation as the spin squeezing generation.
@@
 The squeezing protocol needs a clock laser with a small but nonzero SOC angle $\phi\ll1$, while the clock interrogation protocol requires a clock laser with $\phi=0$.
-Simply using one clock laser with $\phi\ne0$ for the squeezing protocol and another clock laser with $\phi=0$ for the clock interrogation protocol, however, does not resolve this discrepancy, because 
-%(i) both the squeezing and clock interrogation protocols need a clock laser that is homogeneous -- i.e.~with Hamiltonian proportional to $S_\x$,
-%(i) lasers with different SOC angles are homogeneous in different gauges, and 
-%(ii) spin squeezing is a gauge-dependent quantity.
-%More precisely,
-a state that is squeezed with respect to spin operators that are homogeneous (i.e.~of the form $S_\x,S_\y,S_\z$) in a particular gauge is not necessarily squeezed with respect to spin operators that are homogeneous in a different gauge.
+Simply using one clock laser with $\phi\ne0$ for the squeezing protocol and another clock laser with $\phi=0$ for the clock interrogation protocol, however, does not resolve this discrepancy, because a state that is squeezed with respect to spin operators that are homogeneous (i.e.~of the form $S_\x,S_\y,S_\z$) in a particular gauge is not necessarily squeezed with respect to spin operators that are homogeneous in a different gauge.
 In this appendix, we will work explicitly in the ``lab gauge'' of the clock interrogation protocol, in which the Fermi-Hubbard Hamiltonian is SOC-free and the $\phi=0$ clock laser is homogeneous.
@@
 \end{align}
-Defining on-site spin operators in the lab gauge 
+Defining on-site spin operators (in the lab gauge)
 \begin{align}
@@
 If carefully tuning the relative orientations of three clock lasers proves to be too difficult in practice, converting OAT into TAT would therefore need to be done with a pulsed drive protocol, as in Ref.~[\citenum{liu2011spin}].
-% note to authors: the use of a third laser is also incompatible with a 3D accordion lattice, but I (MP) decided that this point would require too much extra discussion to explain here
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Decoherence in the 3D $^{87}$Sr optical lattice clock}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Solving Heisenberg equations of motion for collective spin systems}
@@
 
-\section{ Effect of a harmonic confining trap}
-\label{App:harmonic_trap}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Effect of a harmonic confining trap}
+\label{sec:harmonic_trap}
 \setcounter{figure}{0}
@@
 We briefly review the structure of the single-particle eigenstates of the system, before discussing the effects on squeezing.
-In the quasi-momentum basis, the eigenfunctions $\psi_{n,\alpha}(q) = \braket{q|n,\alpha}$ are given by the $\pi$-periodic Mathieu functions, with the corresponding energies described by the Mathieu characteristic values\cite{Rey2005}.
+In the quasi-momentum basis, the eigenfunctions $\psi_{n,\alpha}(q) = \braket{q|n,\alpha}$ are given by the $\pi$-periodic Mathieu functions, with the corresponding energies described by the Mathieu characteristic values\cite{rey2005ultracold}.
 In the presence of SOC, using  the gauge transformation described in the main text, we obtain the relation
@@
 Concerning the latter point, it is desirable to maintain a weak trap so as to enable a large number of delocalized modes in the trap center, which are the only type capable of contributing to the generation of squeezing.
-Though the interactions between these modes are not strictly all-to-all, they remain long-ranged, and can thus still lead to a spin-locking effect and a protective gap\cite{rey2014probing,Smale2018}.
+Though the interactions between these modes are not strictly all-to-all, they remain long-ranged, and can thus still lead to a spin-locking effect and a protective gap\cite{rey2014probing, smale2019observation}.
 For $\Omega/J = 0.01$ we have $n_c = 28$, enabling $\sim10^3$ contributing modes in each 2D layer of our system.
@@
 which in turn may alter the density distribution and invalidate the spin model.
-Since doublon formation in the localized edges will not have consequences for our squeezing protocol, we must only ensure that doublons are not formed in the trap center, which requires $U > \Omega (n_c/2)^2 = 2J$\cite{Pupillo2006}.
+Since doublon formation in the localized edges will not have consequences for our squeezing protocol, we must only ensure that doublons are not formed in the trap center, which requires $U > \Omega (n_c/2)^2 = 2J$\cite{pupillo2006extended}.
 In Fig.~\ref{fig:harmonic_trap_MB}, we perform exact simulations to assess the effect of the trap on our system.
@@
 \section{Accounting for $p$-wave inelastic collisions}
-\label{App:inelastic_collision}
+\label{sec:inelastic_collision}
 \setcounter{figure}{0}
@@
 \caption{
+{\bf $p$-wave loss rates}.
 Both the averaged $p$-wave inelastic collision rate $\gamma$ (orange) and the ratio of this collision rate to the optimal squeezing rate $\chi_{\text{opt}}$ (blue) are suppressed as the lattice depth increases.
-%The blue line shows  the ratio of the inelastic collision rate $\gamma$ and the optimal squeezing rate $\chi_{\text{opt}}$.
 $\chi_{\text{opt}}$ is obtained by choosing SOC angles $\phi$ that saturate $\widetilde{B}/U\approx0.05$, where $\widetilde{B}$ is the variance of the SOC-induced axial field and $U$ is the two-atom on-site interaction energy.
-%The ratio decreases with increasing lattice depth.
 }
@@
 and compares it with $\gamma=0$ (dashed lines) for  different particle numbers.
-Inelastic collisions prevent the growth of optimal  squeezing with  particle number. For $N=1000$, the maximum squeezing saturates to $\sim 10$dB.
+Inelastic collisions prevent the growth of optimal squeezing with particle number.
+For $N=1000$, the maximum squeezing saturates to $\sim 10$ dB.
 }
@@
 \begin{align}
-\frac{d\hat{\rho}}{dt}&=-\frac{i}{\hbar}[\hat{H}_{\text{eff}},\,\hat{\rho}]+\mathcal{L}\hat{\rho}, 
+  \frac{d\hat{\rho}}{dt}
+  = -\frac{i}{\hbar} [\hat{H}_{\text{eff}},\,\hat{\rho}]
+  + \mathcal{L}\hat{\rho},
 \end{align}
@@
 This latter term can be written using a pseudo-potential approximation as\cite{rey2014probing}:
-
-\begin{eqnarray}
-\mathcal{L}\hat\rho&=&\sum_{{\bm k}{\bm k'} }\Gamma_{{\bm k}{\bm k'} }
-\Big(\hat{A}_{{\bm k}{\bm k'}}\hat\rho\hat{A}_{{\bm k}{\bm k'}}^{\dagger}
--\frac{1}{2}\big(\hat{A}_{{\bm k}{\bm k'}}^{\dagger}\hat{A}_{{\bm k}{\bm k'}}\hat\rho
-+\hat\rho\hat{A}_{{\bm k}{\bm k'}}^{\dagger}\hat{A}_{{\bm k}{\bm k'}}\big)\Big),\nonumber
-\end{eqnarray} where the jump operators are $\hat{A}_{{\bm k}{\bm k'}}=\hat{c}_{{\bm k},\up}\hat{c}_{{\bm k'},\up}$, and
+\begin{multline}
+  \mathcal{L}\hat\rho
+  = \sum_{{\bm k}{\bm k'}} \Gamma_{{\bm k}{\bm k'}}
+  \bigg[\hat{A}_{{\bm k}{\bm k'}}
+  \hat\rho\hat{A}_{{\bm k}{\bm k'}}^\dag \\
+  - \f12 \p{\hat{A}_{{\bm k}{\bm k'}}^\dag \hat{A}_{{\bm k}{\bm k'}} \hat\rho
++\hat\rho\hat{A}_{{\bm k}{\bm k'}}^\dag \hat{A}_{{\bm k}{\bm k'}}}\bigg],
+\end{multline}
+where the jump operators are $\hat{A}_{{\bm k}{\bm k'}}=\hat{c}_{{\bm k},\up}\hat{c}_{{\bm k'},\up}$, and
 $\bm k$, $\bm k'$ sum over all the populated quasi-momentum modes. The decay matrix elements $\Gamma_{{\bm k}{\bm k'}}$ are given by:
 \begin{align}
-\Gamma_{{\bm k}{\bm k'}}&=\frac{3\pi \hbar b_{\text{im}}^3 }{m}\Big(\int d{\bm r}^{\,3}
-W[\phi_{{\bm k}}({\bm r}),\phi_{{\bm k'}}({\bm r})]\Big),\end{align} where $b_{\text{im}} = 121 a_0$\cite{Zhang2014,goban2018emergence} is the $p$-wave inelastic scattering length (with $ a_0=5.29\times 10 ^{-11}$ m the Bohr radius), $\phi_{{\bm k}}({\bm r})$ is the Bloch function with quasi-momentum ${\bm k}$, and $W[\phi_{{\bm k}}({\bm r}),\phi_{{\bm k'}}({\bm r})]\equiv
-[\big({\bm \nabla}\phi^{*}_{{\bm k}}(\bm r)\big)\phi^{*}_{{\bm k'}}(\bm r)
--\phi^{*}_{{\bm k}}(\bm r)\big({\bm \nabla}\phi^{*}_{{\bm k'}}(\bm r)\big)\big]\nonumber\\
-\cdot\big[\big({\bm \nabla}\phi_{{\bm k}}(\bm r)\big)\phi_{{\bm k'}}(\bm r)
--\phi_{{\bm k}}(\bm r)\big({\bm \nabla}\phi_{{\bm k'}}(\bm r)\big)\big]$.
+  \Gamma_{{\bm k}{\bm k'}}
+  = \frac{3\pi \hbar b_{\text{im}}^3 }{m} \p{\int d{\bm r}^{\,3}
+    W[\phi_{{\bm k}}({\bm r}),\phi_{{\bm k'}}({\bm r})]},
+\end{align}
+where $b_{\text{im}} = 121 a_0$\cite{zhang2014spectroscopic, goban2018emergence} is the $p$-wave inelastic scattering length (with $ a_0=5.29\times 10 ^{-11}$ m the Bohr radius), $\phi_{{\bm k}}({\bm r})$ is the Bloch function with quasi-momentum ${\bm k}$, and
+\begin{multline}
+  W\sp{\phi_{{\bm k}}({\bm r}),\phi_{{\bm k'}}({\bm r})} \\
+  \equiv \sp{\p{{\bm \nabla}\phi^{*}_{{\bm k}}(\bm r)}
+  \phi^{*}_{{\bm k'}}(\bm r)-\phi^{*}_{{\bm k}}(\bm r)
+  \p{{\bm \nabla}\phi^{*}_{{\bm k'}}(\bm r)}} \\
+  \cdot\sp{\p{{\bm \nabla}\phi_{{\bm k}}(\bm r)}
+  \phi_{{\bm k'}}(\bm r)-\phi_{{\bm k}}(\bm r)
+  \p{{\bm \nabla}\phi_{{\bm k'}}(\bm r)}}
+\end{multline}
 
-In Fig.~\ref{fig:inelastic_rates} we show  the averaged decay rate $\gamma\equiv\sum_{{\bm k}{\bm k'}}\Gamma_{{\bm k}{\bm k'}}/\ell^2$, where $\ell$ is the number of lattice sites along the $x$ and $y$ axes, as a function of the  lattice depth $V_0$ along these axes.
-Here we assume  the same lattice depth  in the $z$ direction used in the main text, $V=60E_{\text{R}}$.
-As can be seen, the decay rate $\gamma$ is largely suppressed as $V_0$ increases  but is not completely negligible at shallow lattices.
+In Fig.~\ref{fig:inelastic_rates}, we show the averaged decay rate $\gamma\equiv\sum_{{\bm k}{\bm k'}}\Gamma_{{\bm k}{\bm k'}}/\ell^2$, where $\ell$ is the number of lattice sites along the $x$ and $y$ axes, as a function of the lattice depth $V_0$ along these axes.
+Here, we assume the same lattice depth in the $z$ direction used in the main text, $V=60E_{\text{R}}$.
+The decay rate $\gamma$ is suppressed exponentially with increasing lattice depth  $V_0$.
 To quantify the effect of these losses on the spin squeezing generation process, we follow a similar methodology to the one described in detail in Ref.~\cite{rey2014probing}.
-The basic idea is to take advantage of the so-called Truncated-Wigner Approximation (TWA)\cite{Polkovnikok2010, Schachenmayer2015}, which allows us to capture the development of spin squeezing using semi-classical phase-space methods.
+The basic idea is to take advantage of the so-called Truncated-Wigner Approximation (TWA)\cite{polkovnikov2010phase, schachenmayer2015manybody}, which allows us to capture the development of spin squeezing using semi-classical phase-space methods.
 In the TWA the quantum dynamics are accounted for by solving mean field equations of motion supplemented by noise.
 The mean field equations are  derived by assuming that the many-body density matrix of the system can be factorized as
-$\hat{\rho}= \bigotimes_{i} \hat{\rho}({i})$, where $\hat{\rho}(i)$ is the reduced density matrix of the  particle in quasi-momentum  mode  ${\bm q}_i$ (see Eqn.~\eqref{eq:decay_matrix}).
+$\hat{\rho}= \bigotimes_{i} \hat{\rho}({i})$, where $\hat{\rho}(i)$ is the reduced density matrix of the  particle in quasi-momentum  mode  ${\bm q}_i$ [see Eqn.~\eqref{eq:decay_matrix}].
 Under this assumption, the non-linear mean field equations are given by
 \begin{align}
-\frac{d\rho_{\up\up}(j)}{dt}&=-\sum_{j'}\Gamma_{{\bm k}_j{\bm k}_{j'}} \rho_{\up\up}(j) \rho_{\up\up}(j'), \qquad \frac{d\rho_{\dn \dn}(j)}{dt}=0\\
-\frac{d\rho_{\up\dn}(j)}{dt}&= \rho_{\up\dn}(j) \sum_{j'}\Big ( i  \chi(\rho_{\up\up}(j')-\rho_{\dn\dn}(j'))-\frac{\Gamma_{{\bm k}_j{\bm k}_{j'}}}{2} \rho_{\up\up}(j')\Big),\nonumber
-\label{eq:MeanfieldEq}
+  \frac{d\rho_{\up\up}(j)}{dt}
+  = -\sum_{j'}\Gamma_{{\bm k}_j{\bm k}_{j'}}
+  \rho_{\up\up}(j) \rho_{\up\up}(j'),
+  &&
+  \frac{d\rho_{\dn \dn}(j)}{dt}=0
 \end{align}
+and
+\begin{multline}
+  \frac{d\rho_{\up\dn}(j)}{dt}
+  = \rho_{\up\dn}(j) \sum_{j'}
+  \bigg[i\chi(\rho_{\up\up}(j')-\rho_{\dn\dn}(j')) \\
+  - \f12 \Gamma_{{\bm k}_j{\bm k}_{j'}} \rho_{\up\up}(j')\bigg],
+\end{multline}
 where $\rho_{\sigma\sigma'}\equiv\langle\hat{\rho}_{\sigma\sigma'}\rangle$.
@@
 \begin{align}
-\frac{d\rho_{\up\up}^T}{dt}&=-f\gamma  (\rho_{\up\up}^T)^2, \qquad \frac{d\rho_{\dn \dn}^T}{dt}=0\\
-\frac{d\rho_{\up\dn}^T}{dt}&= \rho_{\up\dn}^T \Big (i  \chi(\rho_{\up\up}^T-\rho_{\dn\dn}^T)-\frac{f\gamma}{2} \rho_{\up\up}^T\Big),\nonumber
-\label{eq:MeanfieldEq}
+  \frac{d\rho_{\up\up}^T}{dt} = -f\gamma  (\rho_{\up\up}^T)^2,
+  &&
+  \frac{d\rho_{\dn \dn}^T}{dt} = 0
+\end{align}
+and
+\begin{align}
+  \frac{d\rho_{\up\dn}^T}{dt}
+  = \rho_{\up\dn}^T
+  \sp{i\chi(\rho_{\up\up}^T-\rho_{\dn\dn}^T)
+    - \f12 f\gamma \rho_{\up\up}^T},
 \end{align}
@@
 Using this approach, Fig.~\ref{fig:inelastic_squeezing} shows numerical simulation results of squeezing over time in the presence of inelastic collisions.
-For shallow lattices, the effect of inelastic collisions can limit the spin squeezing to 10 dB.
+For shallow lattices ($V_0\lesssim7E_{\t{R}}$), the effect of inelastic collisions can limit the spin squeezing to $\sim10$ dB.
 Thus, in this regime, losses are as relevant as light scattering.

### 2019-09-20 soc_squeezing/squeezing_paper/prr/transfer_response.tex
pushing (re)submitted version of squeezing paper

\documentclass[preprint,superscriptaddress]{revtex4-2}

%%% linking references
\usepackage{hyperref}
\hypersetup{
  colorlinks=true,
  linkcolor=blue,
  urlcolor=cyan,
}

%%% physics and math packages
\usepackage{physics} % general physics package
\usepackage{amssymb} % math fonts and symbols
\usepackage{bm} % for making math symbols bold
\usepackage{braket} % for nice brackets (e.g. |X>)
\usepackage{mathtools} % for \coloneqq

%%% shorthands to use inside math environments
\renewcommand{\t}{\text} % text in math mode
\newcommand{\f}[2]{\dfrac{#1}{#2}} % shorthand for fractions
\newcommand{\p}[1]{\left(#1\right)} % parenthesis
\renewcommand{\sp}[1]{\left[#1\right]} % square parenthesis
\renewcommand{\set}[1]{\left\{#1\right\}} % square parenthesis
\renewcommand{\c}{\cdot} % inner product
\renewcommand{\d}{\text{d}} % integration measure, e.g. d^3 x
\renewcommand{\v}{\bm} % bold symbols (e.g. for vectors)
\newcommand{\uv}[1]{\hat{\bm #1}} % unit vectors, or vector operators
\newcommand{\bk}{\Braket}
\renewcommand{\ket}{\Ket}
\renewcommand{\bra}{\Bra}

\newcommand{\B}{\mathcal{B}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\G}{\mathcal{G}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\J}{\mathcal{J}}
\renewcommand{\L}{\mathcal{L}}
\renewcommand{\O}{\mathcal{O}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\Q}{\mathcal{Q}}
\renewcommand{\S}{\mathcal{S}}

\newcommand{\z}{\text{z}}
\newcommand{\x}{\text{x}}
\newcommand{\y}{\text{y}}

\newcommand{\up}{\uparrow}
\newcommand{\dn}{\downarrow}

% remove extra white space inside enumerate environments
\usepackage{enumitem}
\setlist[enumerate]{leftmargin=*}

\setlength{\parindent}{0pt}
\setlength{\parskip}{1ex}

% for colored text
\usepackage[dvipsnames]{xcolor}
\newcommand{\blue}[1]{{\color{blue} #1}}
\newcommand{\red}[1]{{\color{red} #1}}
\newcommand{\green}[1]{{\color{ForestGreen} #1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

Dear Editors of Physical Review Research,

Please find attached our revised manuscript, ``Engineering spin squeezing in a 3D optical lattice with interacting spin-orbit-coupled fermions'', for consideration for publication in Physical Review Research.
Below, we address the recommendations and criticisms made by past referees in the peer-review process for Physical Review X, and detail the corresponding changes that we have made to our manuscript.

We were pleased to find that the referee reports were in unanimous agreement about the quality and content of the manuscript, calling it ``well written'' and saying that ``it offers a pathway towards pushing the record level of precision achievable for atomic clocks''.

The most important feedback from previous referees concerned the potential of our manuscript to attract broader interest outside the optical lattice clock community.
We have addressed these concerns in our responses to the referees, as well as corresponding changes to the manuscript.
In short, we believe that developing realistic proposal to achieve a genuine quantum advantage in a world-class sensor unavoidably requires focusing on a specific platform.
Our work is therefore well placed in the context of a broad scientific effort to a achieve a quantum advantage in emergent technologies.
We hope that with these clarifications, our paper can be accepted for publication in Physical Review Research.

Excerpts from the referee reports are written in \blue{blue}, excerpts from the previous version of our manuscript are written in \red{red}, and excerpts from the revised version of our manuscript are written in \green{green}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{First referee}

\blue{The authors replied to my previous report with exhaustive and
  mostly convincing explanations.}

\blue{In the revised version of the manuscript the novelty of their
  results is now clearly explained, which strengthens the impact of
  the work.}

\blue{The reply to my specific questions, e.g.~on the importance of
  the underlying trapping potential and robustness of the achievable
  spin squeezing in a realistic setting, is quite detailed and
  accompanied by significant extensions of the supplementary
  materials. The new version of the manuscript now contains important
  details on the practical implementation of the suggested protocol,
  while keeping an appropriate format for PRX. In light of the
  authors' reply and of the revisions made, I can now fully recommend
  the manuscript for publication.}

We thank the Referee for their positive appreciation of our work, and for the recommendation for publication in PRX.
We appreciate their comment that the revisions we made to our manuscript throughout the review process with PRX ``clearly explained the novelty of our results'' and ``strengthen[ed] the impact of our work''.

We hope that the editors of Physical Review Research agree with this Referee's assessment.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Second referee}
\label{sec:R2}

\begin{enumerate}
\item Concerning:
  \label{pt:R2.scope}

  \blue{The proposal details a protocol to use the weakly interacting,
    weakly spin-orbit coupled regime of fermions in an optical lattice
    to generate spin squeezing for increased precision in the
    state-of-the-art optical lattice clock.  The individual components
    are achievable with current technologies and have all been
    experimentally demonstrated, with many theoretical and
    experimental advancements from the groups led by the senior
    authors, A. M. Rey and J. Ye. The manuscript is in general
    well-written and accessible to the motivated reader, with
    comprehensive appendices.}

  \blue{This work represents current efforts to enhance measurement
    sensitivity through non-classical states and promises considerable
    levels of spin squeezing for the system under consideration. In
    particular, because it focuses so closely on the current
    state-of-the-art optical lattice clock, it offers a pathway
    towards pushing the record level of precision achievable for
    atomic clocks.}

  \blue{At the same time, because the work focuses so closely on this
    particular lattice clock, the experimentalists most interested in
    this work are likely two of the authors and their colleagues. This
    is my biggest concern about the impact of this work: while it does
    open up one new research opportunity, at the immediate level this
    scope seems limited. I ask that the authors address this point in
    full and the more detailed points below.}

  We appreciate the recognition that our manuscript is ``well-written
  and accessible to the motivated reader, with comprehensive
  appendices''.

  Indeed, our work focuses on current state-of-the-art optical lattice
  clocks, since it aims to push current record levels of precision
  achievable with classical sensors.  Without identifying a concrete
  system with world-class sensing capabilities, our effort would
  reduce to a mere proof-of-principle concept which, despite having
  some merit, would lack any practical utility.  Focusing on a
  specific platform is unavoidable for any {\it realistic} and {\it
    practical} proposal to advance the frontier of metrology.  Our
  work therefore fits well into the context of a worldwide scientific
  effort to achieve a quantum advantage in emergent technologies.

  To emphasize the impact of our work in this new paradigm of enhanced
  quantum metrology with correlated quantum states, and thereby to
  further illustrate the relevance of our proposal to a broader
  scientific community, we have rewritten the last paragraph of our
  introduction (left column of page 2), which previously read:

  \red{The intriguing and non-trivial observation reported here is
    that weakly interacting fermions in a lattice with dynamics
    described by the Hubbard model --- which thus obey fermionic
    quantum statistics and interact via purely contact collisions ---
    can feature all-to-all collective spin interactions that naturally
    generate spin-squeezed states in a manner robust to realistic
    experimental imperfections. In the spirit of quantum simulation,
    our work illustrates that short-range interacting systems can be
    used to emulate dynamical magnetic behaviors found in arrays of
    long-range interacting systems (such as arrays of trapped ions[36]
    or cavity QED systems[37, 38] with interactions mediated by
    phonons or photons, respectively), thus enhancing their potential
    as a metrological resource. Although surprising, this result is in
    line with the recent understandings in quantum research that
    explore how the interplay between well understood simple phenomena
    can give rise to emergent new behaviors. In this direction, our
    work emphasizes that the precise control of interactions, which
    becomes possible in 3D optical lattice systems with quantized
    motional degrees of freedom in all three directions, is a key for
    the implementation of the protocols studied here, which may enable
    new quantum technologies.}

  and now reads:

  \green{Despite an abundance of proof-of-principle experiments with
    entangled states[24, 39], so far only the remarkable example of
    LIGO[40, 41] has demonstrated a quantum advantage in a
    state-of-the-art quantum sensing or measurement system.  The new
    generation of 3D optical lattice systems have fully quantized
    motional degrees of freedom[4], allowing for precise control of
    collisional interactions.  We demonstrate how these interactions
    can naturally give rise to metrologically useful correlated
    many-body fermionic states, opening a path to not only generate
    entanglement, but also harness it to achieve a quantum advantage
    in a world-class sensor.  Such an advance will ultimately deliver
    gains to real-world applications including timekeeping,
    navigation, communication, and our understanding of the
    fundamental laws of nature[42].}

  Similarly, we have changed the following sentence in the
  Conclusions:

  \red{To our knowledge, this is the first proposal to use atomic
    collisions for entanglement generation in state-of-the art atomic
    clocks, opening a path for the practical improvement of
    world-leading quantum sensors using correlated many-body fermionic
    states.}

  to:

  \green{To our knowledge, this is the first proposal to use quantum
    correlations in a many-body fermionic system to push
    state-of-the-art quantum sensors beyond the independent-particle
    regime, thereby achieving a genuine quantum advantage.}

  We hope that these changes highlight the broader scope and relevance
  of our work.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\item Concerning:

  \blue{Abstract (L: line)}

  \blue{1. L15: The statement ``This capability exemplifies a new
    paradigm of using driven non-equilibrium systems to overcome
    current limitations in quantum metrology...'' goes unaddressed in
    the main text. The manuscript could be improved if this larger
    impact and its context were made clear, perhaps in the last
    paragraph of the introduction or in the conclusion.}

  We thank the referee for pointing out that our manuscript could be
  improved by following through on the subject of applying driven
  non-equilibrium systems to quantum metrology.  Accordingly, we have
  added the following text to the introduction, preceding the last
  paragraph (left column of page 2):

  \green{This capability mirrors efforts in other settings, such as NV
    centers[36, 37] and trapped ions[38], to enhance quantum metrology
    through the use of driven non-equilibrium phenomena.}

  As the subject of driven non-equilibrium does not play a central
  role in our manuscript, we have also rewritten the last sentence in
  the abstract to de-emphasize this point somewhat, replacing the
  text:

  \red{This capability exemplifies a new paradigm of using driven
    non-equilibrium systems to overcome current limitations in quantum
    metrology, allowing OLCs to enter a new regime of enhanced sensing
    with correlated quantum states.}

  by:

  \green{This capability allows OLCs to enter a new era of quantum
    enhanced sensing using correlated quantum states of driven
    non-equilibrium systems.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\item Concerning:

  \blue{Introduction (P: paragraph, L: line, skipping equations)}

  \blue{1. P4L11: While the orientation of the clock laser and optical
    lattice depth were parameters that were varied to examine the
    resulting optimum level of squeezing, strictly speaking these do
    not need to be actively controlled (beyond pointing/intensity
    stabilization) to generate spin squeezing itself; they just need
    to be set to a single value -- unless I am missing
    something. Perhaps this could be reworded.}

  Indeed, our squeezing protocol does not require {\it active} control
  over clock laser orientation and optical lattice depth.  To clarify
  this point, we have changed the text (page 2, top left paragraph):

  \red{To generate spin squeezing, our protocol only requires control
    over...}

  to:

  \green{To generate spin squeezing, our protocol only requires the
    capability to fix...}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\item Concerning comments about the last paragraph of our introduction
  prior to revision:

  \blue{2. P5L4: The formulation ``...interact via purely contact
    collisions'' is a bit odd. Do you mean, ``solely interact via
    contact collisions''?}

  \blue{3. P5L8: The sentence ``In the spirit of quantum
    simulation... potential as a metrological resource'' seems a bit
    odd to me. Is the point that systems with short-range interactions
    can simulate long-range interacting systems, or that systems with
    short-range interactions can be used in metrology (because of
    effective long-range interactions)? Currently it reads as though
    the second point is being made, in which case I would remove ``In
    the spirit of quantum simulation'' and the parenthetical remark
    because metrology seems to be the main focus. Otherwise I would
    remove ``thus enhancing their potential as a metrological
    resource.''}

  \blue{4. P5L20: ``...which becomes possible in 3D optical lattice
    systems with quantized motional degrees of freedom in all three
    directions...''  As opposed to 3D optical lattices \_without\_
    quantized motional degrees of freedom in all three directions? The
    authors likely mean something along the lines of ``as a result of
    quantized motional degrees of freedom...''}

  These points are no longer applicable, as we have removed the
  relevant paragraph from our manuscript entirely, as detailed in
  point \ref{pt:R2.scope} of our response above.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\item Concerning:

  \blue{From the Fermi-Hubbard model to one-axis and two-axis
    twisting}

  \blue{1. P9L11: In writing ``...up through U/J=8'', it sounds as if
    there is disagreement beyond U/J=8. To prevent this confusion, I
    would reword the statement.}

  To clarify this point, we have changed the text:

  \red{... up through $U/J=8$.}

  to:

  \green{... up through (and exceeding) $U/J=8$.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\item Concerning:

  \blue{2. P9L17: In the sentence beginning with ``Note that we choose
    filling f=5/6...'' it may be good to explicitly state that the
    optimum squeezing occurs at half-filling f=1.}

  To clarify this point, we have changed the sentence:

  \red{Note that we chose filling $f=5/6$ to demonstrate that our
    protocol should work even in this highly hole-doped case; in
    practice, optimized experiments are capable of achieving fillings
    closer to $f=1$.}

  to:

  \green{Note that we chose filling $f=5/6$ to demonstrate that our
    protocol should work, albeit sub-optimally, even in this highly
    hole-doped case; in practice, optimized experiments are capable of
    achieving fillings closer to the optimal $f=1$.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\item Concerning:

  \blue{3. FIG 2: In (b.ii), is it possible to use the same vertical
    axis range as (a.ii)? This would facilitate comparison between the
    two fillings.}

  We have modified figure 2 of the main text accordingly.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\item Concerning:

  \blue{Experimental preparation of a squeezed state and practical
    considerations}

  \blue{1. P1: Rewording this paragraph could be a good opportunity to
    frame the current work beyond the single scope of the
    state-of-the-art OLC.}

  We thank the referee for identifying an opportunity for us to frame
  our work in a broader context.  Accordingly, we have replaced the
  first two sentences of Section III:

  \red{Here we discuss specific implementations of the above protocols
    in the state-of-the-art 3D $^{87}$Sr optical lattice clock (OLC).
    This system has demonstrated the capability to load a quantum
    degenerate gas into a 3D lattice at the ``magic wavelength''
    ($\lambda_{\t{lattice}}=2a\approx813$ nm) for which both the
    ground ($^1S_0,\dn$) and first excited ($^3P_0,\up$) electronic
    states of the atoms experience the same optical potential[4].}

  by:

  \green{Thus far, we have largely considered the general preparation
    of spin-squeezed states with the Fermi-Hubbard model.  Here, we
    discuss the specific implementation of the above protocols in the
    state-of-the-art 3D $^{87}$Sr optical lattice clock (OLC).  If
    successful, such an implementation would (to our knowledge) for
    the first time break through the proof-of-principle stage of spin
    squeezing efforts, and achieve a genuine metrological enhancement
    of a world-class quantum sensor.}

  \green{As required for our protocol, 3D $^{87}$Sr OLC has
    demonstrated the capability to load a quantum degenerate gas into
    a 3D lattice at the ``magic wavelength''
    ($\lambda_{\t{lattice}}=2a\approx813$ nm) for which both the
    ground ($^1S_0,\dn$) and first excited ($^3P_0,\up$) electronic
    states (i.e.~the ``clock states'') of the atoms experience the
    same optical potential[4].}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\item Concerning:

  \blue{2. FIG 3: It may be helpful to the reader to add a dotted line
    to (b) and (c) indicating the boundary where the optical squeezing
    times for OAT and TAT are equal, to emphasize how TAT is faster
    for N$\sim>$500 atoms as noted in the caption.}

  We have added a dotted line to figure 3 marking the boundary at
  which optimal squeezing times for OAT and TAT are equal.
  Accordingly, we have modified the last sentence in the caption of
  figure 3, from:

  \red{... TAT ... achieves this squeezing faster for $N\gtrsim500$
    atoms.}

  to:

  \green{... TAT ... achieves optimal squeezing faster for
    $N\gtrsim400$ atoms, as denoted by a dotted line in panels ({\bf
      b}, {\bf c}).}

  Note that the change from \red{$N\gtrsim500$} to
  \green{$N\gtrsim400$} reflects the boundary between faster optimal
  squeezing under OAT vs.~TAT more accurately.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\item Concerning:

  \blue{3. P4L20: When citing the level of spin squeezing attainable
    with your protocol, including some context or comparison may be
    helpful for the non-expert to evaluate the impact of your
    protocol.}

  We thank the referee for realizing that our manuscript would benefit
  from putting the amount of spin squeezing we report into context
  with other results in the literature.  We have consequently split
  off the following text in Section III:

  \red{The results in Fig.~4 show that squeezing via OAT saturates
    with system size around $N\approx10^3$ ($\ell\approx30$), while
    TAT allows for continued squeezing gains through $N=10^4$
    ($\ell=100$).  Even with decoherence, our protocol may
    realistically generate $\sim10$--$14$ dB of spin squeezing in
    $\sim1$ second with $\sim10^2$--$10^4$ atoms in a 2D section of
    the lattice, which is compatible with the atom numbers and
    interrogation times of state-of-the-art optical lattice clocks[4,
    5].}

  into a separate paragraph that elaborates on levels of spin
  squeezing attained in the literature:

  \green{The results in Fig.~4 show that squeezing via OAT saturates
    with system size around $N\approx10^3$ ($\ell\approx30$), while
    TAT allows for continued squeezing gains through $N=10^4$
    ($\ell=100$).  Even with decoherence, our protocol may
    realistically generate $\sim10$--$14$ dB of spin squeezing in
    $\sim1$ second with $\sim10^2$--$10^4$ atoms in a 2D section of
    the lattice, which is compatible with the atom numbers and
    interrogation times of state-of-the-art optical lattice clocks[4,
    5].  This amount of spin squeezing exceeds those reported in the
    ground-state nuclear spin sublevels of a state-of-the-art
    ${}^{171}$Yb OLC ($\sim6.5$ dB)[55].  While the latter protocol
    might be used to transfer spin squeezing to the electronic clock
    state, to date there has been no demonstration of spin squeezing
    in an optical clock transition.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\item Concerning:

  \blue{4. FIG 4: For this figure, as well as for figures 2 and 3, it
    may be helpful to use different colors and colorbars, one
    color(bar) for squeezing and one for time, kept consistent
    throughout.}

  At the request of the referee, we have made the color maps used in
  figures 2, 3, and 4 all distinct.  We have done so by changing the
  color map in figure 3, so as to keep a larger number of perceptually
  distinct colors in figure 4 (which makes it easier to match a color
  to a numerical value by eye).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\item Concerning:

  \blue{5. FIG 4 caption: It would be helpful to note that the maximum
    value of the colorbar is about 20 dB less than the x-range maximum
    in figure 3a.}

  At the request of the referee, we have added the following sentence
  to the caption of figure 4:

  \green{The decoherence considered in this work also limits maximally
    achievable squeezing to $\sim20$ dB less than the decoherence-free
    maxima shown in Fig.~3.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\item Concerning:

  \blue{6. P5L3: Please quantify what you mean by ``shallow
    lattices'', both here and at the end of Appendix H.}

  This comment concerns words ``shallow lattices'' in the first
  sentence of the text:

  \red{In addition to light scattering, $p$-wave losses from inelastic
    ${}^3 P_0$ collisions[17, 19, 50] can also degrade the maximum
    achievable spin squeezing, particularly in shallow lattices.  More
    details on $p$-wave losses are discussed in Appendix H, where we
    show that operating at lattices deeper than $8 E_{\t{R}}$ may be
    necessary to suppress the impact of inelastic collisions on spin
    squeezing, at the cost of slightly increasing light scattering.}

  The purpose of the first sentence, which contains the words
  ``shallow lattices'', was not to quantify relevant lattice depths,
  but rather to communicate the fact that $p$-wave losses generally
  get worse with increasing lattice depth.  The {\it following}
  sentence then quantifies the implications of $p$-wave losses for
  operating lattice depths.  Our wording previously made this
  structure unclear, so we have changed the above sentences to read:

  \green{In addition to light scattering, $p$-wave losses from
    inelastic ${}^3 P_0$ collisions[17, 19, 56], can also degrade the
    maximum achievable spin squeezing, which becomes more pronounced
    for shallower lattices.  More details on $p$-wave losses are
    discussed in Appendix H, where we show that operating at lattice
    depths $V_0\gtrsim 7 E_{\t{R}}$ may be necessary to suppress the
    impact of inelastic collisions on spin squeezing, at the cost of
    slightly increasing light scattering.}

  Note that the change from \red{$8E_{\t{R}}$} to \green{$7E_{\t{R}}$}
  better reflects our findings in Appendix H, which determined that
  useful squeezing is still achievable in the presence of with
  $p$-wave losses with $U/J\gtrsim6$, at which $V_0\gtrsim7E_{\t{R}}$.

  For similar reasons, we have changed the following text in Appendix
  H:

  \red{As can be seen, the decay rate $\gamma$ is largely suppressed
    as $V_0$ increases but is not completely negligible at shallow
    lattices.  To quantify the effect ...}

  to:

  \green{The decay rate $\gamma$ is suppressed exponentially with
    increasing lattice depth $V_0$.  To quantify the effect ...}

  and we have changed the text (in the last paragraph of Appendix H):

  \red{For shallow lattices, the effect of inelastic collisions can
    limit the spin squeezing to $\sim10$ dB.}

  to:

  \green{For shallow lattices ($V_0\lesssim7E_{\t{R}}$), the effect of
    inelastic collisions can limit the spin squeezing to $\sim10$ dB.}

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{External consultant}

\begin{enumerate}
\item Concerning:

  \blue{The present manuscript is a theory proposal motivated very
    much by the existing, and highly optimized setup of a Sr lattice
    clock in the Ye lab at JILA. There is no question that this is a
    world leading setup. To push the accuracy further, one question is
    how -- in this existing setup -- one can generate spin squeezed
    states. I note that spin squeezing is probably relevant for clocks
    involving a relatively small number of atoms (like present trapped
    ion experiments), but for lattice clocks with a very large number
    the relevance is a question to ask. Let me also mention that spin
    squeezing has been demonstrated with cavities etc. (Thompson at
    JILA, Vuletic MIT and others), but its actual usefulness in an
    actual leading atomic clock still awaits demonstration, as far as
    I know.}

  \blue{What this paper does is to start from Hubbard models, and
    design One-Axis-Twist Hamiltonians (there is a huge literature on
    this) from the Hubbard Hamiltonian naturally appearing in the
    lattice clock in the limit of shallow lattices. This is probably a
    limit they find it necessary to work in because of other
    constraints. Their idea is to get an infinite-range interaction
    required for One-Axis-Twisting by inducing an energy gap so that
    Hubbard model reduces to the One-Axis-Twist
    Hamiltonian. Conceptual ideas like this is what I have seen
    before, e.g.~in
    \url{https://journals.aps.org/pra/abstract/10.1103/PhysRevA.80.032311},
    though in a different experimental context.}

  As the consultant says, spin squeezing has indeed been demonstrated
  in other experimental platforms, and despite numerous proposals such
  as
  \href{https://journals.aps.org/pra/abstract/10.1103/PhysRevA.80.032311}{PhysRevA.80.032311},
  the ``actual usefulness'' of spin squeezing still awaits
  demonstration.  In our view, this point highlights the relevance and
  importance of our work, which goes beyond proof-of-principle
  demonstrations that spin squeezing might be used to achieve some
  technological advantage.  Specifically, our work outlines the first
  (to our knowledge) {\it realistic} proposal to generate {\it useful}
  spin squeezing that would push the state-of-the-art in a world-class
  sensor.  To emphasize this point, we have rewritten the last
  paragraph of our introduction (as detailed in point
  \ref{pt:R2.scope} of Section \ref{sec:R2} above).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\item Concerning:

  \blue{My conclusions are:}

  \blue{Pushing atomic clocks, and lattice clocks in particular, to a
    next level of accuracy, and exploring the use of spin squeezing,
    and in particular its use in an actual precision experiment, is
    highly relevant.}

  \blue{The present proposal might be a good way to go in the Jun Ye
    lab, and would thus have quite an impact if successful.}

  \blue{However, this paper is quite specialized to the JILA setup,
    and the question is if these are ideas which are conceptually new,
    and applicable in a broader context. Here my answer would be
    slightly mixed: using Spins Orbit Coupling to open gaps in the
    spectrum to make infinite range interactions is an interesting
    idea. But is it enough?}

  \blue{Of course, if they can demonstrate the existence of squeezing
    based on these ideas in a real experiment at JILA on whatever
    level, I would be unreservedly impressed and enthusiastic. As a
    theory paper alone, it might be a bit too specialized. It is the
    old question and dilemma for theorists: publish a theory paper by
    itself, or wait until experimentalists actually do it -- which
    might take a long time.}

  We appreciate the recognition that our proposal is ``highly
  relevant'', with the potential to have ``quite an impact''.  In
  light of this recognition, and the admission that the consultant
  would be ``unreservedly impressed and enthusiastic'' with a
  successful experimental implementation of our proposal, with all due
  respect we must express firm disagreement with the consultant's
  present takeaway from the ``old question and dilemma'' of whether
  theorists should publish their ideas, or wait until these ideas are
  implemented experimentally.  Due to the scarcity of realistic
  proposals for a practical quantum advantage in world-class sensors,
  we believe that our proposal and methods are worthwhile
  disseminating to a broad scientific audience that can scrutinize our
  ideas, and build on them to develop their own (possibly modified or
  improved) implementations.  We hope that the editors agree with our
  view on the importance to widely disseminate not only experimental
  results, but also theoretical proposals that guide experiments in
  ways that actually advance state-of-the-art sensors and pave the
  ground for the practical development and application of new
  technologies.

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Miscellaneous changes}

\begin{enumerate}
\item To clarify our high-level summary of the mechanism for spin
  squeezing, we have added the following sentence to the summary of
  our results in the introduction (second to last paragraph of Section
  I, top left of page 2 in the revised text):

  \green{Interactions thereby prolong inter-particle spin coherence
    through a spin-locking effect, which additionally transforms the
    dephasing effect of SOC into a collective spin squeezing process.}

\item To add structure to our manuscript, we have changed the title of
  Section II from ``\red{From the Fermi-Hubbard model to one-axis and
    two-axis twisting}'' to ``\green{Spin squeezing with the
    Fermi-Hubbard model}'', and we have added the sub-section headings
  ``\green{Model validity}'' (II.A) and ``\green{Two-axis twisting}''
  (II.B).

\item We have shortened the title of Section III, from
  ``\red{Experimental preparation of a squeezed state and practical
    considerations}'' to ``\green{Experimental implementation and
    practical considerations}''.

\item To clarify some notation, we have changed the last sentence
  before the conclusion, from:

  \red{Such a pulse can transfer $\ket{g, I=-7/2}$ to
    $\ket{e, I=-9/2}$ without affecting $\ket{g, I=-9/2}$.}

  to:

  \green{Such a pulse can transfer $\ket{g,-7/2}$ to $\ket{e,-9/2}$
    without affecting $\ket{g,-9/2}$, where $g$ and $e$ respectively
    denote the ground and excited (electronic) clock states.}

\item We have added the following text to the beginning of the caption
  to figure H1:

  \green{{\bf $p$-wave loss rates}.}
\end{enumerate}

\end{document}

### 2019-09-20 soc_squeezing/squeezing_paper/supplementary.tex
pushing (re)submitted version of squeezing paper

--- 2019-05-21 soc_squeezing/squeezing_paper/supplementary.tex
@@
-\documentclass[prx,superscriptaddress,12pt]{revtex4-2}
+\documentclass[aps,prx,superscriptaddress,notitlepage,onecolumn,12pt]{revtex4-1}
 
@@
 
-%%% for referencing files in the main text
-\usepackage{xr}
-\externaldocument[M-]{main}
-
 \begin{document}
@@
 \author{P.~He} \contrib
-\author{M.A.~Perlin} \contrib
-\author{S.R.~Muleady}
-\author{R.J.~Lewis-Swan}
+\author{M.~A.~Perlin} \contrib
+\author{S.~R.~Muleady}
+\author{R.~J.~Lewis-Swan}
 \affiliation{\JILA}
 \affiliation{\CTQM}
-\author{R.B.~Hutson}
+\author{R.~B.~Hutson}
 \author{J.~Ye}
 \affiliation{\JILA}
-\author{A.M.~Rey}
+\author{A.~M.~Rey}
 \affiliation{\JILA}
@@
 
-In principle, spin-changing decoherence compromises the validity of the OAT model, as its perturbative derivation in Appendix \ref{M-sec:derivation_OAT} of the main text relies on spin population remaining primarily within the Dicke manifold.
+In principle, spin-changing decoherence compromises the validity of the OAT model, as its perturbative derivation in Appendix A of the main text relies on spin population remaining primarily within the Dicke manifold.
 This assumption breaks down in the presence of, for example, spontaneous emission, which transfers population outside of the Dicke manifold.
@@
 In the absence of decoherence, OAT achieves a maximal amount of squeezing that scales as $\xi^2\sim N^{-2/3}$, while TAT achieves Heisenberg-limited squeezing with $\xi^2\sim N^{-1}$.
-Note that our method for computing squeezing via TAT in the presence of decoherence (described in Appendix \ref{M-sec:collective_simulation} of the main text) is not capable of computing squeezing for the full range of times shown in Fig.~\ref{fig:squeezing_example}; the corresponding time-series data in this figure is therefore shown up to the point at which this method breaks down.
+Note that our method for computing squeezing via TAT in the presence of decoherence (described in Appendix F of the main text) is not capable of computing squeezing for the full range of times shown in Fig.~\ref{fig:squeezing_example}; the corresponding time-series data in this figure is therefore shown up to the point at which this method breaks down.
 
@@
 Atoms are confined along the direction transverse to the 2D layer by a lattice of depth 60 $E_{\t{R}}$.
-Squeezing over time is shown for OAT (blue) and TAT (green), both with (solid lines) and without (dashed lines) decoherence via uncorrelated decay and dephasing of individual spins at rates of $0.1~\t{sec}^{-1}$ (see Appendix \ref{M-sec:decoherence} of the main text).
+Squeezing over time is shown for OAT (blue) and TAT (green), both with (solid lines) and without (dashed lines) decoherence via uncorrelated decay and dephasing of individual spins at rates of $0.1~\t{sec}^{-1}$ (see Appendix E of the main text).
 }

### 2019-09-20 soc_squeezing/scripts/plot_benchmarking.m

@@ -117,7 +117,10 @@ function [ acceptable_params ] = ...
     ylabel_obj.Units = 'normalized';
     ylabel_obj.Position = ylabel_obj.Position + [0 0.02 0];
     if strcmp(dtype, 't')
-        z_max = max(max(data_max));
+        %%% WARNING: setting axis limits manually
+        %z_max = max(max(data_max));
+        z_max = 13;
+        %%%
         set(gca, 'ZLim', [0 z_max], 'ZTick', 0:5:z_max);
         text(0.1, 1/10, z_max*1.05, '$\times10^2$', 'interpreter', 'latex');
     else % if dtype == 'sq'

### 2019-09-20 soc_squeezing/scripts/plot_squeezing_optima.py

@@ -31,7 +31,6 @@ depths_dec = np.arange(20,81,2)/10 # lattice depths in plots
 sizes_dec = np.arange(10,101,5) # linear lattice sizes in plots
 phi_cutoff = 10 # minimum value of (\phi/\pi)^{-1}
 
-color_map = "jet" # for the 2-D color plots
 dpi = 600
 
 params = { "font.family" : "sans-serif",
@@ -103,7 +102,7 @@ for method in sqz_methods:
     sqz_opt = -to_dB(pd_read_1D(data_dir + f"sqz_{method}.txt"))
     t_opt[method] = pd_read_2D(data_dir + f"optimization/t_opt_{method}_L_2D.txt")
     t_opt[method] /= recoil_energy_NU
-    depths, sizes = t_opt[method].index, t_opt[method].columns
+    depths, sizes = t_opt[method].index, np.array(t_opt[method].columns)
 
     err = abs(depths[1] - depths[0]) / 2
     use_depths = (depths >= depth_min-err) & (depths <= depth_max+err)
@@ -116,15 +115,21 @@ sqz_max = max([ sqz_vals[method].max() for method in sqz_methods ])
 t_min = min([ t_opt[method].values.min() for method in sqz_methods ])
 t_max = max([ t_opt[method].values.max() for method in sqz_methods ])
 
+faster_NOAT = t_opt[NOAT] > t_opt[OAT]
+
 # plot data
 for method in sqz_methods:
     ax_sqz.plot(sqz_vals[method], sizes, lines[method], label = method)
 
     data = t_opt[method].values[use_depths,:].T
     mesh = ax[method].pcolormesh(depths[use_depths], sizes, data,
-                                 cmap = plt.get_cmap(color_map),
-                                 vmin = t_min, vmax = t_max,
+                                 cmap = plt.get_cmap("twilight_shifted"),
+                                 vmin = 0, vmax = t_max,
                                  zorder = 0, rasterized = True)
+
+    ax[method].contour(depths[use_depths], sizes,
+                       faster_NOAT.values[use_depths,:].T,
+                       levels = 0, colors = [ "k" ], linestyles = "dotted")
 
     ax[method].set_xlabel(depth_label, zorder = 1)
     ax[method].set_yticklabels([])
@@ -250,7 +255,7 @@ fig, ( ax[OAT], ax[NOAT], cax ) \
 # plot data
 for method in sqz_methods:
     mesh = ax[method].pcolormesh(depths[method], sizes[method], sqz[method].T,
[…]

### 2019-09-24 sun_squeezing/sun_notes.tex
adding some notes about multi-level generalizations of collective spin dynamics and squeezing

\documentclass[nofootinbib,notitlepage,11pt]{revtex4-2}

%%% linking references
\usepackage{hyperref}
\hypersetup{
  breaklinks=true,
  colorlinks=true,
  linkcolor=blue,
  filecolor=magenta,
  urlcolor=cyan,
}

%%% header / footer
\usepackage{fancyhdr} % easier header and footer management
\pagestyle{fancy} % page formatting style
\fancyhf{} % clear all header and footer text
\renewcommand{\headrulewidth}{0pt} % remove horizontal line in header
\usepackage{lastpage} % for referencing last page
\cfoot{\thepage~of \pageref{LastPage}} % "x of y" page labeling

%%% symbols, notations, etc.
\usepackage{physics,braket,bm,amssymb} % physics and math
\renewcommand{\t}{\text} % text in math mode
\newcommand{\f}[2]{\dfrac{#1}{#2}} % shorthand for fractions
\newcommand{\p}[1]{\left(#1\right)} % parenthesis
\renewcommand{\sp}[1]{\left[#1\right]} % square parenthesis
\renewcommand{\set}[1]{\left\{#1\right\}} % curly parenthesis
\newcommand{\bk}{\Braket} % shorthand for braket notation
\renewcommand{\v}{\bm} % bold vectors
\newcommand{\uv}[1]{\bm{\hat{#1}}} % unit vectors
\newcommand{\av}{\vec} % arrow vectors
\renewcommand{\d}{\text{d}} % for infinitesimals
\renewcommand{\c}{\cdot} % inner product

\usepackage{dsfont} % for identity operator
\newcommand{\1}{\mathds{1}}

\newcommand{\x}{\text{x}}
\newcommand{\y}{\text{y}}
\newcommand{\z}{\text{z}}

\newcommand{\D}{\mathcal{D}}
\renewcommand{\O}{\mathcal{O}}
\newcommand{\T}{\mathcal{T}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\title{Multi-level generalizations of collective spin dynamics and
  squeezing}%
\author{Michael A. Perlin}%
\date{\today}

\maketitle
\thispagestyle{fancy}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Uniform interactions and the Generalized Gell-Mann matrices}

Under the frozen-mode approximation, SU($n$)-symmetric uniform
all-to-all two-body interactions between fermions on a lattice take
the form\cite{perlin2019effective}
\begin{align}
  H_{\t{int}} = \f{u}{2} \sum_{p,q,\mu,\nu}
  \p{c_{p\mu}^\dag c_{q\nu}^\dag c_{q\nu} c_{p\mu}
    + c_{q\mu}^\dag c_{p\nu}^\dag c_{q\nu} c_{p\mu}},
\end{align}
where $u\equiv U/L$ is the two-body on-site interaction energy $U$
divided by the number of lattice sites $L$, and $c_{p\mu}$ is the
annihilation operator for a fermionic mode with quasi-momentum $p$ and
internal state (e.g.~nuclear spin) $\mu$.  Defining the spin operators
$s_{\mu\nu}^{(p)}\equiv c_{p\mu}^\dag c_{p\nu}$ for each
quasi-momentum $p$, we can alternately write
\begin{align}
  H_{\t{int}} = \f{u}{2} \sum_{p,q,\mu,\nu}
  \p{s_{\mu\mu}^{(p)} s_{\nu\nu}^{(q)} - s_{\mu\nu}^{(p)} s_{\nu\mu}^{(q)}}
  = \f{u}{2}\p{N^2 - \sum_{p,q,\mu,\nu} s_{\mu\nu}^{(p)} s_{\nu\mu}^{(q)}},
\end{align}
where $N$ is the total number of particles.  In the absence of
coherence between sectors of different particle number, we can neglect
the $\sim N^2$ term and simply write
\begin{align}
  H_{\t{int}}
  = - \f{u}{2} \sum_{p,q,\mu,\nu} s_{\mu\nu}^{(p)} s_{\nu\mu}^{(q)}.
  \label{eq:H_int_start}
\end{align}
In the case of SU(2), at this point we would expand the spin operators
$s_{\mu\nu}^{(p)}$ in terms of Pauli operators in order to convert
$H_{\t{int}}$ into an SU(2) spin Hamiltonian.  To generalize this
procedure to SU($n$), we define the {\it generalized Gell-Mann (GGM)
  operators}\cite{hioe1981level, bertlmann2008bloch} for
$j,k,\ell\in\set{0,1,\cdots,n-1}$ with $k<j$ and $\ell>0$:
\begin{align}
  \lambda_0 \equiv \sqrt{\f{2}{n}}~ \1,
  &&
  \lambda_\ell \equiv \sqrt{\f{2}{\ell\p{\ell+1}}}
  \p{\sum_{n=0}^{\ell-1}\op{n} - \ell \op{\ell}},
  \label{eq:GGM_diag}
\end{align}
\begin{align}
  \lambda_{jk,+} \equiv \op{j}{k} + \op{k}{j},
  &&
  \lambda_{jk,-} \equiv i\p{\op{j}{k} - \op{k}{j}},
  \label{eq:GGM_off_diag}
\end{align}
where $\1$ is the identity operator.  The GGM operators in
\eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag} coincide exactly with
the Pauli operators in the case of $n=2$, generally provide a complete
basis for operators on the Hilbert space of an $n$-level system, and
satisfy the orthonormality relation
\begin{align}
  \bk{\lambda_a,\lambda_b}
  \equiv \tr\p{\lambda_a^\dag\lambda_b}
  = 2\delta_{ab}.
  \label{eq:GGM_inner}
\end{align}
Denoting a vector of all GGM operators by $\v\lambda$, we can make the
remarkable simplification
\begin{align}
  H_{\t{int}}
  = -\f{u}{4} \sum_{p,q} \v\lambda^{(p)} \c \v\lambda^{(q)}
  = -u \v\Lambda \c \v\Lambda,
  \label{eq:H_int_GGM}
\end{align}
where $\O^{(p)}$ denotes the action of $\O$ on the degrees of freedom
at quasi-momentum mode $p$, and
$\v\Lambda\equiv\f12\sum_p\v\lambda^{(p)}$ is a collective SU($n$)
spin vector.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Transition operators}

The GGM operators in \eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag}
provide a convenient operator basis to describe dynamics obeying
SU($n$) symmetry.  An external driving field addressing $n$-level
nuclear spins, however, will generally violate SU($n$) symmetry.  Such
a field obeys the symmetries of the {\it polarization} or {\it
  transition operators} defined by\cite{kryszewski2006alternative,
  bertlmann2008bloch}
\begin{align}
  T_{JM} \equiv \sqrt{\f{2J+1}{2I+1}} \sum_{\mu,\nu}
  \bk{I\mu;I\nu,JM} \op{\mu}{\nu},
  \label{eq:trans_ops}
\end{align}
where $J\in\set{0,1,\cdots,n-1}$ and $M\in\set{-J,-J+1,J}$ index a
total spin and its projection onto a quantization axis;
$I\equiv\p{n-1}/2$ is the maximal angular momentum of a spin-$n$
system; $\mu,\nu\in\set{-I,-I+1,\cdots,I}$ index projections of an
$n$-level nuclear spin onto a quantization axis; and
$\bk{I\mu,JM;I\nu}$ is a Clebsch-Gordan coefficient.  Similarly to the
GGM operators in \eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag}, the
transition operators in \eqref{eq:trans_ops} provide a complete basis
for the space of operators on the Hilbert space of an $n$-level
system, and satisfy the orthonormality relation
\begin{align}
  \bk{T_{JM},T_{J'M'}}
  \equiv \tr\p{T_{JM}^\dag T_{J'M'}}
  = \delta_{JJ'} \delta_{MM'}.
\end{align}
Written in terms of the transition operators, the interaction
Hamiltonian in \eqref{eq:H_int_start} takes the form
\begin{align}
  H_{\t{int}} = -\f{u}{2} \sum_{p,q} {\v T^{(p)}}^\dag \c \v T^{(q)}
  = -\f{u}{2} \v\T^\dag \c \v\T,
  \label{eq:H_int_trans}
\end{align}
where $\v T$ is a vector of all transition operators in
\eqref{eq:trans_ops} and $\v\T\equiv\sum_p\v T^{(p)}$ is a collective
transition vector.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Drive operators}

Unlike the GGM operators in \eqref{eq:GGM_diag} and
\eqref{eq:GGM_off_diag}, the transition operators in
\eqref{eq:trans_ops} are not self-adjoint.  In order to express
Hamiltonians in terms of self-adjoint operators, we define the {\it
  drive operators}
\begin{align}
  D_{J,0} \equiv \sqrt{2}~ T_{J,0},
  &&
  D_{J\Delta,+} \equiv -\p{T_{J\Delta} + T_{J\Delta}^\dag},
  &&
  D_{J\Delta,-} \equiv i\p{T_{J\Delta} - T_{J\Delta}^\dag},
  \label{eq:drive_ops}
\end{align}
for $\Delta\in\set{1,2,\cdots,J}$.  The drive operators satisfy the
same orthonormality relations as the GGM operators, provided in
\eqref{eq:GGM_inner}, and are similarly equal to the Pauli operators
in the case of $n=2$.  These operators have the convenient feature
that $D_{1,1,\pm}$ is (up to a gauge transformation) proportional to
the Hamiltonian induced on a spin-$n$ system by a classical driving
laser.  In terms of the drive operators,
\begin{align}
  H_{\t{int}} = -\f{u}{4} \sum_{p,q} \v D^{(p)}\c\v D^{(q)}
  = -u \v\D \c \v\D,
  \label{eq:H_int_drive}
\end{align}
where $\v D$ is a vector of all drive operators in
\eqref{eq:drive_ops} and $\v\D\equiv\f12\sum_p\v D^{(p)}$ is a
collective drive vector.

\bibliography{sun_notes.bib}

\end{document}

### 2019-09-25 sun_squeezing/sun_notes.tex
added a proof that the collective interaction Hamiltonian takes an identical form in any orthonormal basis of operators added a comment about Weyl operators

@@
 
+\newcommand{\B}{\mathcal{B}}
 \newcommand{\D}{\mathcal{D}}
+\newcommand{\I}{\mathcal{I}}
+\newcommand{\N}{\mathcal{N}}
 \renewcommand{\O}{\mathcal{O}}
+\newcommand{\Q}{\mathcal{Q}}
 \newcommand{\T}{\mathcal{T}}
+\renewcommand{\S}{\mathcal{S}}
+
+\def\obra#1{\mathinner{({#1}|}}
+\def\oket#1{\mathinner{|{#1})}}
+\def\obk#1{\mathinner{({#1})}}
+\def\oop#1#2{\oket{#1}\!\obra{#2}}
 
@@
 \begin{document}
+\thispagestyle{fancy}
 
@@
 \maketitle
-\thispagestyle{fancy}
+
+\tableofcontents
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Uniform interactions and the Generalized Gell-Mann matrices}
+\section{Collective SU($n$)-symmetric interactions}
 
-Under the frozen-mode approximation, SU($n$)-symmetric uniform
-all-to-all two-body interactions between fermions on a lattice take
-the form\cite{perlin2019effective}
+Under the frozen-mode approximation, two-body collective
+SU($n$)-symmetric interactions between fermions on a lattice take the
+form\cite{perlin2019effective}
 \begin{align}
@@
   = - \f{u}{2} \sum_{p,q,\mu,\nu} s_{\mu\nu}^{(p)} s_{\nu\mu}^{(q)}.
-  \label{eq:H_int_start}
 \end{align}
@@
 $H_{\t{int}}$ into an SU(2) spin Hamiltonian.  To generalize this
-procedure to SU($n$), we define the {\it generalized Gell-Mann (GGM)
-  operators}\cite{hioe1981level, bertlmann2008bloch} for
-$j,k,\ell\in\set{0,1,\cdots,n-1}$ with $k<j$ and $\ell>0$:
+procedure to the case of SU($n$), we first define a vector
+$\v s^{(p)}$ of all spin operators $s_{\mu\nu}^{(p)}$, and write
+\begin{align}
+  H_{\t{int}}
+  = - \f{u}{2} \sum_{p,q} {\v s^{(p)}}^\dag \v s^{(q)}
+  = - \f{u}{2} \v\S^\dag \c \v\S,
+  &&
+  \v\S\equiv \sum_p \v s^{(p)}.
+  \label{eq:H_int_spin}
+\end{align}
+In the following sections, we identify various bases for the space of
+operators on the Hilbert space of an $n$-level system.  For brevity,
+we denote this space of operators by $\B_n$, and note that $\B_n$ is
+itself an $n^2$-dimensional Hilbert space equipped with a trace
+(Hilbert-Schmidt) inner product
+\begin{align}
+  \obk{\O|\Q} \equiv \tr\p{\O^\dag Q},
+\end{align}
+where $\oket{\Q}$ and $\obra{\O}$ respectively denote vectors in
+$\B_n$ and its dual space $\B_n^*$.  Similarly to the set of spin
+operators $\set{s_{\mu\nu}}$ on the Hilbert space at each
+quasi-momentum, all of the bases we consider will be orthonormal with
+respect to the trace inner product.  As we show in Appendix
+\ref{sec:changing_bases}, the collective SU($n$)-symmetric interaction
+Hamiltonian in \eqref{eq:H_int_spin} takes an identical form in any
+orthonormal basis for $\B_n$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Generalized Gell-Mann (GGM) operators}
+
+The most obvious generalization of Pauli operators to the case of
+SU($n$) are the {\it generalized Gell-Mann (GGM)
+  operators}\cite{hioe1981level, bertlmann2008bloch}, defined for
+$j,k,\ell\in\set{0,1,\cdots,n-1}$ with $k<j$ and $\ell>0$ by
 \begin{align}
@@
 \end{align}
-where $\1$ is the identity operator.  The GGM operators in
-\eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag} coincide exactly with
-the Pauli operators in the case of $n=2$, generally provide a complete
-basis for operators on the Hilbert space of an $n$-level system, and
-satisfy the orthonormality relation
+where $\1$ is the identity operator.  The self-adjoint GGM operators
+in \eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag} provide a complete
+basis for the space of operators on the Hilbert space of an $n$-level
+system, coincide exactly with the Pauli operators in the case of
+$n=2$, and satisfy the orthonormality condition
 \begin{align}
-  \bk{\lambda_a,\lambda_b}
-  \equiv \tr\p{\lambda_a^\dag\lambda_b}
-  = 2\delta_{ab}.
+  \obk{\lambda_a|\lambda_b} = 2\delta_{ab}.
   \label{eq:GGM_inner}
 \end{align}
-Denoting a vector of all GGM operators by $\v\lambda$, we can make the
-remarkable simplification
+Denoting a vector of all GGM operators by $\v\lambda$, we can
+therefore write (see Appendix \ref{sec:changing_bases})
 \begin{align}
@@
   = -u \v\Lambda \c \v\Lambda,
+  &&
+  \v\Lambda \equiv \f12 \sum_p \v\lambda^{(p)}.
   \label{eq:H_int_GGM}
 \end{align}
-where $\O^{(p)}$ denotes the action of $\O$ on the degrees of freedom
-at quasi-momentum mode $p$, and
-$\v\Lambda\equiv\f12\sum_p\v\lambda^{(p)}$ is a collective SU($n$)
-spin vector.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Transition operators}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Transition operators}
 
@@
 SU($n$) symmetry.  An external driving field addressing $n$-level
-nuclear spins, however, will generally violate SU($n$) symmetry.  Such
-a field obeys the symmetries of the {\it polarization} or {\it
+nuclear spins, however, will generally violate SU($n$) symmetry, and
+instead obeys the symmetries of the {\it polarization} or {\it
   transition operators} defined by\cite{kryszewski2006alternative,
@@
 for the space of operators on the Hilbert space of an $n$-level
-system, and satisfy the orthonormality relation
+system, and satisfy the orthonormality condition
 \begin{align}
-  \bk{T_{JM},T_{J'M'}}
-  \equiv \tr\p{T_{JM}^\dag T_{J'M'}}
-  = \delta_{JJ'} \delta_{MM'}.
+  \obk{T_{JM}|T_{J'M'}} = \delta_{JJ'} \delta_{MM'},
 \end{align}
-Written in terms of the transition operators, the interaction
-Hamiltonian in \eqref{eq:H_int_start} takes the form
+which implies
 \begin{align}
@@
   = -\f{u}{2} \v\T^\dag \c \v\T,
+  &&
+  \v\T \equiv \v\T\equiv\sum_p\v T^{(p)},
   \label{eq:H_int_trans}
@@
 where $\v T$ is a vector of all transition operators in
-\eqref{eq:trans_ops} and $\v\T\equiv\sum_p\v T^{(p)}$ is a collective
-transition vector.
+\eqref{eq:trans_ops}.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Drive operators}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Drive operators}
 
@@
 \end{align}
-for $\Delta\in\set{1,2,\cdots,J}$.  The drive operators satisfy the
-same orthonormality relations as the GGM operators, provided in
-\eqref{eq:GGM_inner}, and are similarly equal to the Pauli operators
-in the case of $n=2$.  These operators have the convenient feature
-that $D_{1,1,\pm}$ is (up to a gauge transformation) proportional to
-the Hamiltonian induced on a spin-$n$ system by a classical driving
-laser.  In terms of the drive operators,
+for $\Delta\in\set{1,2,\cdots,J}$.  Similarly to the GGM operators in
+\eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag}, the drive operators
+satisfy the orthonormality condition
+\begin{align}
+  \obk{D_a|D_b} = 2\delta_{ab},
+\end{align}
+and are similarly equal to the Pauli operators in the case of $n=2$.
+These operators have the convenient feature that $D_{1,1,\pm}$ is (up
+to a gauge transformation) proportional to the Hamiltonian induced on
+a spin-$n$ system by a classical driving laser.  In terms of the drive
+operators,
 \begin{align}
@@
   = -u \v\D \c \v\D,
+  &&
+  \v\D \equiv \f12 \sum_p \v D^{(p)},
   \label{eq:H_int_drive}
@@
 where $\v D$ is a vector of all drive operators in
-\eqref{eq:drive_ops} and $\v\D\equiv\f12\sum_p\v D^{(p)}$ is a
-collective drive vector.
+\eqref{eq:drive_ops}.
+
+We note that Weyl operators\cite{bertlmann2008bloch} and self-adjoint
+linear combinations thereof\cite{asadian2016heisenbergweyl} also
+provide an interesting basis of operators that generalize the notion
+of position and momentum to the case of a finite-dimensional space.
+As our present work is motivated by multi-level systems that are
+realized with nuclear spin (angular momentum) states, however, the
+transition and drive operators are more natural to consider than the
+Weyl operators, so we will not discuss the latter any further.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{External drive}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Spin-orbit coupling}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Classifying eigenstates}
+
+\appendix
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Changing operator bases for collective SU($n$)-symmetric
+  interactions}
+\label{sec:changing_bases}
+
+Here we show that the collective interaction Hamiltonian in
+\eqref{eq:H_int_spin} takes an identical form in any orthonormal basis
+for $\B_n$, i.e.~the space of operators on the Hilbert space of an
+$n$-level system, equipped with the trace (Hilbert-Schmidt) inner
+product
+\begin{align}
+  \obk{\O|\Q} \equiv \tr\p{\O^\dag \Q}.
+\end{align}
+To maintain generality, we choose an arbitrary basis $\set{X_j}$
+satisfying the orthonormality condition
+\begin{align}
+  \obk{X_j|X_k} = \N_X \delta_{jk},
+\end{align}
+for $j\in\set{0,1,\cdots,n^2-1}$.  This orthonormality condition
+implies that we can resolve the identity (super-)operator $\I$ with
+$\I\oket{\O}=\oket{\O}$ for any $\oket\O\in\B_n$ as
+\begin{align}
+  \I = \f1{\N_X} \sum_j \oop{X_j}{X_j}
+  = \f1{\N_X} \sum_j \oop{X_j^\dag}{X_j^\dag},
+\end{align}
+where we used the fact that $\set{X_j^\dag}$ is also an orthonormal
+basis for $\B_n$, satisfying the same orthonormality condition as
+$\set{X_j}$; these two bases are transformed into each other by
+unitary $\sum_j\oop{X_j^\dag}{X_j}/\sqrt{\N_X}$.
+
+There are two types of terms in the Hamiltonian in
+\eqref{eq:H_int_spin}: two-body interactions and self-interactions,
+respectively of the form
+\begin{align}
+  \sum_j X_j^\dag \otimes X_j
+  = \sum_j \oket{X_j^\dag}\otimes\oket{X_j},
+  &&
+  \sum_j X_j^\dag X_j = \sum_j \oket{X_j^\dag X_j}.
+\end{align}
+Choosing a different basis $\set{Y_j}$ for $\B_n$ with norm $\N_Y$, by
+resolution of the identity we find that
+\begin{align}
+  \sum_j \oket{X_j^\dag}\otimes\oket{X_j}
+  = \f1{\N_Y^2} \sum_{j,k,\ell}
+  \oket{Y_k^\dag} \obk{Y_k^\dag|X_j^\dag}
+  \otimes \oket{Y_\ell} \obk{Y_\ell|X_j},
+\end{align}
+where
+\begin{align}
+  \obk{Y_k^\dag|X_j^\dag}
+  = \tr\p{Y_k X_j^\dag}
+  = \tr\p{X_j^\dag Y_k}
+  = \obk{X_j|Y_k},
+\end{align}
+which allows us to simplify
+\begin{align}
+  \sum_j \oket{X_j^\dag}\otimes\oket{X_j}
+  = \f1{\N_Y^2} \sum_{k,\ell}
+  \oket{Y_k^\dag} \otimes \oket{Y_\ell}
+  \sum_j \obk{Y_\ell|X_j} \obk{X_j|Y_k}
+  = \f{\N_X}{\N_Y} \sum_k
+  \oket{Y_k^\dag} \otimes \oket{Y_k}.
+\end{align}
+For the single-body self-interactions, we can similarly expand
+\begin{align}
+  X_j = \f1{\N_Y} \sum_k Y_k \obk{Y_k|X_j},
+  &&
+  X_j^\dag = \f1{\N_Y} \sum_k Y_k^\dag \obk{Y_k^\dag|X_j^\dag}
+  = \f1{\N_Y} \sum_k Y_k^\dag \obk{X_j|Y_k},
+\end{align}
+which implies
+\begin{align}
+  \sum_j X_j^\dag X_j
+  = \f1{\N_Y^2} \sum_{j,k,\ell} Y_k^\dag Y_\ell
+  \obk{X_j|Y_k} \obk{Y_\ell|X_j}
+  = \f{\N_X}{\N_Y} \sum_k Y_k^\dag Y_k.
+\end{align}
+The collective interaction Hamiltonian in \eqref{eq:H_int_spin} thus
+takes an identical form in any orthonormal basis for the space $\B_n$
+of operators on the Hilbert space of an $n$-level system, up to
+appropriate re-scaling factors $\N_X/\N_Y$.
 

### 2019-09-25 sun_squeezing/sun_notes.tex
simplified derivation for changing operator bases in collective interactions

@@
 
-There are two types of terms in the Hamiltonian in
-\eqref{eq:H_int_spin}: two-body interactions and self-interactions,
-respectively of the form
+The collective interaction Hamiltonian in \eqref{eq:H_int_spin} takes
+the form
 \begin{align}
-  \sum_j X_j^\dag \otimes X_j
-  = \sum_j \oket{X_j^\dag}\otimes\oket{X_j},
+  H = \sum_{p,q,j} {X_j^{(p)}}^\dag X_j^{(q)},
+\end{align}
+which essentially consists of two-body interactions ($p\ne q$) and
+self-interactions ($p=q$), respectively of the form
+\begin{align}
+  \sum_j X_j \otimes X_j^\dag,
   &&
-  \sum_j X_j^\dag X_j = \sum_j \oket{X_j^\dag X_j}.
+  \sum_j X_j X_j^\dag.
 \end{align}
-Choosing a different basis $\set{Y_j}$ for $\B_n$ with norm $\N_Y$, by
-resolution of the identity we find that
-\begin{align}
-  \sum_j \oket{X_j^\dag}\otimes\oket{X_j}
-  = \f1{\N_Y^2} \sum_{j,k,\ell}
-  \oket{Y_k^\dag} \obk{Y_k^\dag|X_j^\dag}
-  \otimes \oket{Y_\ell} \obk{Y_\ell|X_j},
-\end{align}
-where
-\begin{align}
-  \obk{Y_k^\dag|X_j^\dag}
-  = \tr\p{Y_k X_j^\dag}
-  = \tr\p{X_j^\dag Y_k}
-  = \obk{X_j|Y_k},
-\end{align}
-which allows us to simplify
-\begin{align}
-  \sum_j \oket{X_j^\dag}\otimes\oket{X_j}
-  = \f1{\N_Y^2} \sum_{k,\ell}
-  \oket{Y_k^\dag} \otimes \oket{Y_\ell}
-  \sum_j \obk{Y_\ell|X_j} \obk{X_j|Y_k}
-  = \f{\N_X}{\N_Y} \sum_k
-  \oket{Y_k^\dag} \otimes \oket{Y_k}.
-\end{align}
-For the single-body self-interactions, we can similarly expand
+Choosing a different basis $\set{Y_j}$ for $\B_n$ with corresponding
+norm $\N_Y$, we can expand
 \begin{align}
@@
 \end{align}
-which implies
+where we used the fact that
 \begin{align}
-  \sum_j X_j^\dag X_j
-  = \f1{\N_Y^2} \sum_{j,k,\ell} Y_k^\dag Y_\ell
-  \obk{X_j|Y_k} \obk{Y_\ell|X_j}
-  = \f{\N_X}{\N_Y} \sum_k Y_k^\dag Y_k.
+  \obk{Y_k^\dag|X_j^\dag}
+  = \tr\p{Y_k X_j^\dag}
+  = \tr\p{X_j^\dag Y_k}
+  = \obk{X_j|Y_k}.
+\end{align}
+By resolution of the identity, we can therefore find
+\begin{align}
+  \sum_j X_j \otimes X_j^\dag
+  = \f1{\N_Y^2} \sum_{j,k,\ell} Y_k \obk{Y_k|X_j}
+  \otimes Y_\ell^\dag \obk{X_j|Y_\ell}
+  = \f{\N_X}{\N_Y^2} \sum_{k,\ell} Y_k \otimes Y_\ell^\dag
+  \obk{Y_k|\I|Y_\ell}
+  = \f{\N_X}{\N_Y} \sum_k Y_k \otimes Y_k^\dag,
+\end{align}
+and similarly
+\begin{align}
+  \sum_j X_j X_j^\dag
+  = \f1{\N_Y^2} \sum_{j,k,\ell} Y_k Y_\ell^\dag
+  \obk{Y_k|X_j} \obk{X_j|Y_\ell}
+  = \f{\N_X}{\N_Y} \sum_k Y_k Y_k^\dag.
 \end{align}

### 2019-09-25 sun_squeezing/sun_notes.tex
fixed a minor error and improved some explanatory text

@@
 \begin{align}
-  \obk{\lambda_a|\lambda_b} = 2\delta_{ab}.
+  \obk{\lambda_a|\lambda_b} = 2\delta_{ab},
   \label{eq:GGM_inner}
 \end{align}
-Denoting a vector of all GGM operators by $\v\lambda$, we can
-therefore write (see Appendix \ref{sec:changing_bases})
+where $a,b$ index any GGM operator in \eqref{eq:GGM_diag} or
+\eqref{eq:GGM_off_diag}.  Denoting a vector of all GGM operators by
+$\v\lambda$, we can therefore write (see Appendix
+\ref{sec:changing_bases})
 \begin{align}
@@
 $n$-level nuclear spin onto a quantization axis; and
-$\bk{I\mu,JM;I\nu}$ is a Clebsch-Gordan coefficient.  Similarly to the
+$\bk{I\mu;I\nu,JM}$ is a Clebsch-Gordan coefficient.  Similarly to the
 GGM operators in \eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag}, the
@@
 \eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag}, the drive operators
-satisfy the orthonormality condition
+are equal to the Pauli operators in the case of $n=2$, and satisfy the
+orthonormality condition
 \begin{align}
@@
 \end{align}
-and are similarly equal to the Pauli operators in the case of $n=2$.
-These operators have the convenient feature that $D_{1,1,\pm}$ is (up
-to a gauge transformation) proportional to the Hamiltonian induced on
-a spin-$n$ system by a classical driving laser.  In terms of the drive
+where $a,b$ index any drive operator in \eqref{eq:drive_ops}.  These
+operators have the convenient feature that $D_{1,1,\pm}$ is (up to a
+gauge transformation) proportional to the Hamiltonian induced on a
+spin-$n$ system by a classical driving laser.  In terms of the drive
 operators,
@@
 linear combinations thereof\cite{asadian2016heisenbergweyl} also
-provide an interesting basis of operators that generalize the notion
-of position and momentum to the case of a finite-dimensional space.
-As our present work is motivated by multi-level systems that are
-realized with nuclear spin (angular momentum) states, however, the
+provide an interesting basis of operators, generalizing the notion of
+position and momentum to the case of a finite-dimensional Hilbert
+space.  As our present work is motivated by multi-level systems that
+are realized with nuclear spin (angular momentum) states, however, the
 transition and drive operators are more natural to consider than the

### 2019-09-26 sun_squeezing/sun_notes.tex
added text about an external laser drive and spin-orbit coupling

@@
 
+\newcommand{\up}{\uparrow}
+\newcommand{\dn}{\downarrow}
+
 \newcommand{\x}{\text{x}}
@@
 
-Under the frozen-mode approximation, two-body collective
-SU($n$)-symmetric interactions between fermions on a lattice take the
-form\cite{perlin2019effective}
+We consider a collection of $n$-level fermions on a 1-D lattice, with
+a single-particle Hamiltonian
 \begin{align}
-  H_{\t{int}} = \f{u}{2} \sum_{p,q,\mu,\nu}
+  H_{\t{lat}} = -J \sum_{q,\mu} \cos\p{qa} c_{q\mu}^\dag c_{q\mu},
+  \label{eq:H_lat}
+\end{align}
+where $J$ is the nearest-neighbor tunneling rate, $a$ is the spacing
+between neighboring lattice sites, and $c_{q\mu}$ is the fermionic
+annihilation operator for a particle in quasi-momentum mode $q$ with
+internal state (e.g.~nuclear spin) $\mu$.  Under the frozen-mode
+approximation, two-body collective SU($n$)-symmetric interactions
+between these fermions take form\cite{perlin2019effective}
+\begin{align}
+  H_{\t{int}} = \f12 u \sum_{p,q,\mu,\nu}
   \p{c_{p\mu}^\dag c_{q\nu}^\dag c_{q\nu} c_{p\mu}
@@
 where $u\equiv U/L$ is the two-body on-site interaction energy $U$
-divided by the number of lattice sites $L$, and $c_{p\mu}$ is the
-annihilation operator for a fermionic mode with quasi-momentum $p$ and
-internal state (e.g.~nuclear spin) $\mu$.  Defining the spin operators
-$s_{\mu\nu}^{(p)}\equiv c_{p\mu}^\dag c_{p\nu}$ for each
+divided by the number of lattice sites $L$.  Defining the spin
+operators $s_{\mu\nu}^{(p)}\equiv c_{p\mu}^\dag c_{p\nu}$ for each
 quasi-momentum $p$, we can alternately write
 \begin{align}
-  H_{\t{int}} = \f{u}{2} \sum_{p,q,\mu,\nu}
+  H_{\t{int}} = \f12 u \sum_{p,q,\mu,\nu}
   \p{s_{\mu\mu}^{(p)} s_{\nu\nu}^{(q)} - s_{\mu\nu}^{(p)} s_{\nu\mu}^{(q)}}
-  = \f{u}{2}\p{N^2 - \sum_{p,q,\mu,\nu} s_{\mu\nu}^{(p)} s_{\nu\mu}^{(q)}},
+  = \f12 u\p{N^2 - \sum_{p,q,\mu,\nu} s_{\mu\nu}^{(p)} s_{\nu\mu}^{(q)}},
 \end{align}
@@
   H_{\t{int}}
-  = - \f{u}{2} \sum_{p,q,\mu,\nu} s_{\mu\nu}^{(p)} s_{\nu\mu}^{(q)}.
+  = - \f12 u \sum_{p,q,\mu,\nu} s_{\mu\nu}^{(p)} s_{\nu\mu}^{(q)}.
 \end{align}
@@
   H_{\t{int}}
-  = - \f{u}{2} \sum_{p,q} {\v s^{(p)}}^\dag \v s^{(q)}
-  = - \f{u}{2} \v\S^\dag \c \v\S,
+  = - \f12 u \sum_{p,q} {\v s^{(p)}}^\dag \v s^{(q)}
+  = - \f12 u \v\S^\dag \c \v\S,
   &&
@@
 where $\oket{\Q}$ and $\obra{\O}$ respectively denote vectors in
-$\B_n$ and its dual space $\B_n^*$.  Similarly to the set of spin
-operators $\set{s_{\mu\nu}}$ on the Hilbert space at each
-quasi-momentum, all of the bases we consider will be orthonormal with
-respect to the trace inner product.  As we show in Appendix
-\ref{sec:changing_bases}, the collective SU($n$)-symmetric interaction
-Hamiltonian in \eqref{eq:H_int_spin} takes an identical form in any
-orthonormal basis for $\B_n$.
+$\B_n$ and its dual space $\B_n^*$.  Similarly to the set of all spin
+operators $s_{\mu\nu}\equiv\op{\mu}{\nu}$ on the single-particle
+Hilbert space at each quasi-momentum, all of the bases we consider
+will be orthonormal with respect to the trace inner product.  As we
+show in Appendix \ref{sec:changing_bases}, the collective
+SU($n$)-symmetric interaction Hamiltonian in \eqref{eq:H_int_spin}
+takes an identical form in any orthonormal basis for $\B_n$.
 
@@
   H_{\t{int}}
-  = -\f{u}{4} \sum_{p,q} \v\lambda^{(p)} \c \v\lambda^{(q)}
+  = -\f14 u \sum_{p,q} \v\lambda^{(p)} \c \v\lambda^{(q)}
   = -u \v\Lambda \c \v\Lambda,
@@
 total spin and its projection onto a quantization axis;
-$I\equiv\p{n-1}/2$ is the maximal angular momentum of a spin-$n$
-system; $\mu,\nu\in\set{-I,-I+1,\cdots,I}$ index projections of an
+$I\equiv\p{n-1}/2$ is the maximal angular momentum of an $n$-level
+spin; $\mu,\nu\in\set{-I,-I+1,\cdots,I}$ index projections of an
 $n$-level nuclear spin onto a quantization axis; and
@@
 \begin{align}
-  H_{\t{int}} = -\f{u}{2} \sum_{p,q} {\v T^{(p)}}^\dag \c \v T^{(q)}
-  = -\f{u}{2} \v\T^\dag \c \v\T,
+  H_{\t{int}} = -\f12 u \sum_{p,q} {\v T^{(p)}}^\dag \c \v T^{(q)}
+  = -\f12 u \v\T^\dag \c \v\T,
   &&
@@
 operators have the convenient feature that $D_{1,1,\pm}$ is (up to a
-gauge transformation) proportional to the Hamiltonian induced on a
-spin-$n$ system by a classical driving laser.  In terms of the drive
+gauge transformation) proportional to the Hamiltonian induced on an
+$n$-level spin by a classical driving laser.  In terms of the drive
 operators,
 \begin{align}
-  H_{\t{int}} = -\f{u}{4} \sum_{p,q} \v D^{(p)}\c\v D^{(q)}
+  H_{\t{int}} = -\f14 u \sum_{p,q} \v D^{(p)}\c\v D^{(q)}
   = -u \v\D \c \v\D,
@@
 space.  As our present work is motivated by multi-level systems that
-are realized with nuclear spin (angular momentum) states, however, the
-transition and drive operators are more natural to consider than the
-Weyl operators, so we will not discuss the latter any further.
+are realized with nuclear spin (angular momentum) degrees of freedom,
+however, the transition and drive operators are more natural to
+consider than the Weyl operators, so we will not discuss the latter
+any further.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{External drive}
+\section{External drive and spin-orbit coupling}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Spin-orbit coupling}
+An external driving laser addressing a spin-$1/2$ particle on a
+lattice will induce the Hamiltonian
+\begin{align}
+  \left. H_{\t{drive}}^{(\phi)} \right|_{n=2}
+  = \f12 \Omega \sum_j e^{i\phi_j} c_{j,\up}^\dag c_{j,\dn} + \t{h.c.},
+  \label{eq:H_drive_2}
+\end{align}
+where $\Omega$ is a drive strength, $j$ indexes a lattice site, and
+$\phi_j$ is a site-dependent phase.  The multi-level analogue of
+\eqref{eq:H_drive_2} is
+\begin{align}
+  H_{\t{drive}}^{(\phi)}
+  = -\f12 \Omega \sum_j e^{i\phi_j} T_{1,1}^{(j)} + \t{h.c.}.
+  \label{eq:H_drive}
+\end{align}
+We can make this drive spatially homogenous via the gauge
+transformation $c_{j\mu} \to e^{i\mu\phi_j} c_{j\mu}$, after which the
+drive becomes
+\begin{align}
+  H_{\t{drive}}
+  = -\f12 \Omega \sum_j T_{1,1}^{(j)} + \t{h.c.}
+  = \f12 \Omega \sum_j D_{1,1,+}
+  = \Omega \D_{1,1,+},
+\end{align}
+where $\D_a\equiv\sum_jD_a^{(j)}/2$ is a collective drive operator,
+with e.g.~$\D_{1,1,+}=S_\x$ in the case of $n=2$.  The SU($n$)
+symmetry of collective interactions means that the interaction
+Hamiltonian $H_{\t{int}}$ in \eqref{eq:H_int_drive} remains unchanged
+under this gauge transformation, but the single-particle Hamiltonian
+in \eqref{eq:H_lat} transformed as
+\begin{align}
+  H_{\t{lat}}
+  \to H_{\t{lat}}^{(\phi)}
+  = -J \sum_{q,\mu} \cos\p{qa+\mu\phi} c_{q\mu}^\dag c_{q\mu}
+  = -J \sum_{q,\mu} \cos\p{qa+\mu\phi} s_{\mu\mu}^{(q)},
+\end{align}
+where $\phi$ is the relative phase between adjacent lattice sites.  In
+terms of the drive operators, we can expand
+\begin{align}
+  H_{\t{lat}}^{(\phi)}
+  = -\sum_q \sp{\cos\p{qa} \sum_{J\,\t{even}} B_J D_{J,0}^{(q)}
+    - \sin\p{qa} \sum_{J\,\t{odd}} B_J D_{J,0}^{(q)}},
+\end{align}
+where the effective ``fields'' $B_J$ take the form
+\begin{align}
+  \left. B_J \right|_{J\,\t{even}}
+  = \f12 \sum_\mu \obk{D_{J,0}|s_{\mu\mu}} \cos\p{\mu\phi},
+  &&
+  \left. B_J \right|_{J\,\t{odd}}
+  = \f12 \sum_\mu \obk{D_{J,0}|s_{\mu\mu}} \sin\p{\mu\phi}.
+\end{align}
 
@@
 \section{Classifying eigenstates}
+
+[not yet written]
 

### 2019-09-26 collective_sun/sun_notes.tex
added explicit derivation of SOC hamiltonian in drive operator basis

--- 2019-09-26 sun_squeezing/sun_notes.tex
@@
 \begin{align}
-  T_{JM} \equiv \sqrt{\f{2J+1}{2I+1}} \sum_{\mu,\nu}
-  \bk{I\mu;I\nu,JM} \op{\mu}{\nu},
+  T_{LM} \equiv \sqrt{\f{2L+1}{2I+1}} \sum_{\mu,\nu}
+  \bk{I\mu;I\nu,LM} \op{\mu}{\nu},
   \label{eq:trans_ops}
 \end{align}
-where $J\in\set{0,1,\cdots,n-1}$ and $M\in\set{-J,-J+1,J}$ index a
+where $L\in\set{0,1,\cdots,n-1}$ and $M\in\set{-L,-L+1,L}$ index a
 total spin and its projection onto a quantization axis;
@@
 $n$-level nuclear spin onto a quantization axis; and
-$\bk{I\mu;I\nu,JM}$ is a Clebsch-Gordan coefficient.  Similarly to the
+$\bk{I\mu;I\nu,LM}$ is a Clebsch-Gordan coefficient.  Similarly to the
 GGM operators in \eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag}, the
@@
 \begin{align}
-  \obk{T_{JM}|T_{J'M'}} = \delta_{JJ'} \delta_{MM'},
+  \obk{T_{LM}|T_{L'M'}} = \delta_{LL'} \delta_{MM'},
 \end{align}
@@
 \begin{align}
-  D_{J,0} \equiv \sqrt{2}~ T_{J,0},
+  D_{L,0} \equiv \sqrt{2}~ T_{L,0},
   &&
-  D_{J\Delta,+} \equiv -\p{T_{J\Delta} + T_{J\Delta}^\dag},
+  D_{LM} \stackrel{M>0}{\equiv} -\p{T_{L\Delta} + T_{L\Delta}^\dag},
   &&
-  D_{J\Delta,-} \equiv i\p{T_{J\Delta} - T_{J\Delta}^\dag},
+  D_{LM} \stackrel{M<0}{\equiv} i\p{T_{L\Delta} - T_{L\Delta}^\dag},
   \label{eq:drive_ops}
 \end{align}
-for $\Delta\in\set{1,2,\cdots,J}$.  Similarly to the GGM operators in
-\eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag}, the drive operators
-are equal to the Pauli operators in the case of $n=2$, and satisfy the
-orthonormality condition
+for integer $M$ with $\abs{M}\le L$.  Similarly to the GGM operators
+in \eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag}, the drive
+operators are equal to the Pauli operators in the case of $n=2$, and
+satisfy the orthonormality condition
 \begin{align}
-  \obk{D_a|D_b} = 2\delta_{ab},
+  \obk{D_{LM}|D_{L'M'}} = 2\delta_{LL'}\delta_{MM'},
 \end{align}
 where $a,b$ index any drive operator in \eqref{eq:drive_ops}.  These
-operators have the convenient feature that $D_{1,1,\pm}$ is (up to a
-gauge transformation) proportional to the Hamiltonian induced on an
+operators have the convenient feature that $D_{1,1}$ is (up to a gauge
+transformation) proportional to the Hamiltonian induced on an
 $n$-level spin by a classical driving laser.  In terms of the drive
@@
 linear combinations thereof\cite{asadian2016heisenbergweyl} also
-provide an interesting basis of operators, generalizing the notion of
-position and momentum to the case of a finite-dimensional Hilbert
-space.  As our present work is motivated by multi-level systems that
-are realized with nuclear spin (angular momentum) degrees of freedom,
-however, the transition and drive operators are more natural to
-consider than the Weyl operators, so we will not discuss the latter
-any further.
+provide an interesting basis of operators, generalizing continuous
+phase-space displacement operators to the case of a finite-dimensional
+Hilbert space.  As our present work is motivated by multi-level
+systems that are realized with nuclear spin (angular momentum) degrees
+of freedom, however, the transition and drive operators are more
+natural to consider than the Weyl operators, so we will not discuss
+the latter any further.
 
@@
   = -\f12 \Omega \sum_j T_{1,1}^{(j)} + \t{h.c.}
-  = \f12 \Omega \sum_j D_{1,1,+}
-  = \Omega \D_{1,1,+},
+  = \f12 \Omega \sum_j D_{1,1}
+  = \Omega \D_{1,1},
 \end{align}
-where $\D_a\equiv\sum_jD_a^{(j)}/2$ is a collective drive operator,
-with e.g.~$\D_{1,1,+}=S_\x$ in the case of $n=2$.  The SU($n$)
-symmetry of collective interactions means that the interaction
-Hamiltonian $H_{\t{int}}$ in \eqref{eq:H_int_drive} remains unchanged
-under this gauge transformation, but the single-particle Hamiltonian
-in \eqref{eq:H_lat} transformed as
+where $\D_{LM}\equiv\sum_jD_{LM}^{(j)}/2$ is a collective drive
+operator, with e.g.~$\D_{1,1}=S_\x$ in the case of $n=2$.  The SU($n$)
+symmetry of collective interactions implies that the interaction
+Hamiltonian $H_{\t{int}}$ in \eqref{eq:H_int_drive} is invariant under
+this gauge transformation, but the single-particle Hamiltonian in
+\eqref{eq:H_lat} transforms as
 \begin{align}
@@
   \to H_{\t{lat}}^{(\phi)}
-  = -J \sum_{q,\mu} \cos\p{qa+\mu\phi} c_{q\mu}^\dag c_{q\mu}
-  = -J \sum_{q,\mu} \cos\p{qa+\mu\phi} s_{\mu\mu}^{(q)},
+  = -J \sum_{q,\mu} \cos\p{qa+\mu\phi} c_{q\mu}^\dag c_{q\mu},
+  \label{eq:H_lat_SOC}
 \end{align}
-where $\phi$ is the relative phase between adjacent lattice sites.  In
-terms of the drive operators, we can expand
+where $\phi$ is the relative phase of the driving laser between
+adjacent lattice sites.  In the basis of drive operators, this
+Hamiltonian takes the form (see Appendix \ref{sec:lat_drive})
 \begin{align}
-  H_{\t{lat}}^{(\phi)}
-  = -\sum_q \sp{\cos\p{qa} \sum_{J\,\t{even}} B_J D_{J,0}^{(q)}
-    - \sin\p{qa} \sum_{J\,\t{odd}} B_J D_{J,0}^{(q)}},
+  H_{\t{lat}}^{(\phi)} = -J \sum_{q,L} B_L\p{q,\phi} D_{L,0}^{(q)},
+  \label{eq:H_lat_drive}
 \end{align}
-where the effective ``fields'' $B_J$ take the form
+with effective driving field strengths
 \begin{align}
-  \left. B_J \right|_{J\,\t{even}}
-  = \f12 \sum_\mu \obk{D_{J,0}|s_{\mu\mu}} \cos\p{\mu\phi},
+  B_L\p{q,\phi} = \p{-1}^L f_L\p{qa} \sqrt{\f{L+1/2}{n}}
+  \sum_\mu \bk{I\mu;I\mu,L,0} f_L\p{\mu\phi},
   &&
-  \left. B_J \right|_{J\,\t{odd}}
-  = \f12 \sum_\mu \obk{D_{J,0}|s_{\mu\mu}} \sin\p{\mu\phi}.
+  f_L\p{\theta} \equiv
+  \begin{cases}
+    \cos\theta & L~\t{even} \\
+    \sin\theta & L~\t{odd}
+  \end{cases}.
 \end{align}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Spin-orbit coupling with drive operators}
+\label{sec:lat_drive}
+
+Here we take the single-particle spin-orbit coupled Hamiltonian
+$H_{\t{lat}}^{(\phi)}$ in \eqref{eq:H_lat_SOC}, and derive its
+expression in \eqref{eq:H_lat_drive} in terms of the drive operators
+$D_{LM}$.  In the basis of ordinary spin operators $s_{\mu\mu}$, this
+Hamiltonian takes the form
+\begin{align}
+  H_{\t{lat}}^{(\phi)}
+  = -J \sum_{q,\mu} \cos\p{qa+\mu\phi} s_{\mu\mu}^{(q)}
+  \label{eq:H_lat_spin}
+\end{align}
+Defining the diagonal ``spin-wave'' operators
+\begin{align}
+  \omega_{\phi,+} \equiv \sum_\mu \cos\p{\mu\phi} s_{\mu\mu},
+  &&
+  \omega_{\phi,-} \equiv \sum_\mu \sin\p{\mu\phi} s_{\mu\mu},
+\end{align}
+which are respectively even and odd under spin inversion
+($\mu\to-\mu$), we can simplify
+\begin{align}
+  \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
+  = \cos\p{qa} \omega_{\phi,+} - \sin\p{qa} \omega_{\phi,-},
+\end{align}
+and resolve the identity (super-)operator $\I$ with
+$\I \omega_{\phi,\pm} = \omega_{\phi,\pm}$ in the basis of the drive
+operators, finding
+\begin{align}
+  \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
+  = \cos\p{qa} \f12 \sum_{L,M} \obk{\omega_{\phi,+}|D_{LM}} D_{LM}
+  - \sin\p{qa} \f12 \sum_{L,M} \obk{\omega_{\phi,-}|D_{LM}} D_{LM}.
+\end{align}
+We now note that the drive operators $D_{LM}$ are strictly
+off-diagonal for $M\ne0$, and furthermore that $D_{L,0}$ with even
+(odd) $L$ is even (odd) under spin inversion, which implies
+\begin{align}
+  \sum_{L,M} \obk{\omega_{\phi,+}|D_{LM}} D_{LM}
+  = \sum_{L\,\t{even}} \obk{\omega_{\phi,+}|D_{L,0}} D_{L,0},
+\end{align}
+and likewise with $\omega_{\phi,-}$ and odd $L$.  In total, we can
+write
+\begin{align}
+  \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
+  = \cos\p{qa} \f12 \sum_{L\,\t{even}}
+  \obk{\omega_{\phi,+}|D_{L,0}} D_{L,0}
+  - \sin\p{qa} \f12 \sum_{L\,\t{odd}}
+  \obk{\omega_{\phi,-}|D_{L,0}} D_{L,0},
+\end{align}
+or, more compactly,
+\begin{align}
+  \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
+  = \sum_L B_L\p{q,\phi} D_{L,0},
+  \label{eq:compact_SOC_drive}
+\end{align}
+with effective driving field strengths
+\begin{align}
+  B_L\p{q,\phi}
+  &\equiv \p{-1}^L f_L\p{qa} \f12
+  \sum_\mu f_L\p{\mu\phi} \obk{s_{\mu\mu}|D_{L,0}} \\
+  &= \p{-1}^L f_L\p{qa} \sqrt{\f{L+1/2}{n}}
+  \sum_\mu \bk{I\mu;I\mu,L,0} f_L\p{\mu\phi},
+\end{align}
+where
+\begin{align}
+  f_L\p{\theta} \equiv
+  \begin{cases}
+    \cos\theta & L~\t{even} \\
+    \sin\theta & L~\t{odd}
+  \end{cases}.
+\end{align}
+Substituting the result in \eqref{eq:compact_SOC_drive} into the
+single-particle Hamiltonian in \eqref{eq:H_lat_spin} yields
+\begin{align}
+  H_{\t{lat}}^{(\phi)} = -J \sum_{q,L} B_L\p{q,\phi} D_{L,0}^{(q)}.
+\end{align}
+
 \bibliography{sun_notes.bib}

### 2019-09-26 collective_sun/sun_notes.tex
fixed minor error in appendix on changing operator bases

@@
   \left. H_{\t{drive}}^{(\phi)} \right|_{n=2}
-  = \f12 \Omega \sum_j e^{i\phi_j} c_{j,\up}^\dag c_{j,\dn} + \t{h.c.},
+  = \f12 \Omega \sum_j
+  \p{e^{i\phi_j} c_{j,\up}^\dag c_{j,\dn}
+    + e^{-i\phi_j} c_{j,\dn}^\dag c_{j,\up}},
   \label{eq:H_drive_2}
@@
   H_{\t{drive}}^{(\phi)}
-  = -\f12 \Omega \sum_j e^{i\phi_j} T_{1,1}^{(j)} + \t{h.c.}.
+  = -\f12 \Omega \sum_j \p{e^{i\phi_j} T_{1,1}^{(j)}
+    + e^{-i\phi_j} {T_{1,1}^{(j)}}^\dag}.
   \label{eq:H_drive}
@@
   H_{\t{drive}}
-  = -\f12 \Omega \sum_j T_{1,1}^{(j)} + \t{h.c.}
+  = -\f12 \Omega \sum_j\p{T_{1,1}^{(j)} + {T_{1,1}^{(j)}}^\dag}
   = \f12 \Omega \sum_j D_{1,1}
@@
 $\set{X_j}$; these two bases are transformed into each other by
-unitary $\sum_j\oop{X_j^\dag}{X_j}/\sqrt{\N_X}$.
+unitary $\sum_j\oop{X_j^\dag}{X_j}/\N_X$.
 
@@
 \begin{align}
-  H = \sum_{p,q,j} {X_j^{(p)}}^\dag X_j^{(q)},
+  H = \sum_{p,q,j} X_j^{(p)} {X_j^{(q)}}^\dag,
 \end{align}

### 2019-09-27 collective_sun/sun_notes.tex
added discussion mapping drive operators onto spin operators of a collective SU(2) system

@@
   &&
-  \v\T \equiv \v\T\equiv\sum_p\v T^{(p)},
+  \v\T \equiv \sum_p\v T^{(p)},
   \label{eq:H_int_trans}
@@
 \end{align}
-where $a,b$ index any drive operator in \eqref{eq:drive_ops}.  These
-operators have the convenient feature that $D_{1,1}$ is (up to a gauge
-transformation) proportional to the Hamiltonian induced on an
-$n$-level spin by a classical driving laser.  In terms of the drive
-operators,
+where $a,b$ index any drive operator in \eqref{eq:drive_ops}.  In
+terms of the drive operators,
 \begin{align}
@@
 \eqref{eq:drive_ops}.
+
+The drive operators have the convenient feature that $D_{1,0}$ and
+$D_{1,1}$ are (up to a gauge transformation) respectively proportional
+to the Hamiltonians induced on an $n$-level spin by an external
+magnetic field and a classical driving laser; loosely speaking, one
+should identify $\p{D_{1,0},D_{1,1}}\sim\p{S_\z,S_\x}$, where
+$S_\alpha$ is a collective SU(2) spin-$\alpha$
+operator\cite{perlin2019shorttime}.  More concretely, we can embed the
+Hilbert space of a single $n$-level spin into the
+permutationally-symmetric (Dicke) manifold of states for $\p{n-1}$
+2-level spins via
+\begin{align}
+  \ket{\mu}_{\t{single}} \to \ket{\mu+I}_{\t{collective}}
+  \propto S_+ \ket\dn^{\otimes\p{n-1}},
+\end{align}
+where $I\equiv\p{n-1}/2$ is a total spin,
+$\mu\in\set{-I,-I+1,\cdots,I}$ indexes a state of the single $n$-level
+spin, and $S_+$ is a collective spin-raising operator for $\p{n-1}$
+2-level spins.  This embedding takes
+\begin{align}
+  \p{D_{1,0},D_{1,1},D_{1,-1}}
+  \to \p{S_\z,S_\x,S_\y}
+  \times \sqrt{\f{3}{I \p{I+1/2} \p{I+1}}}.
+\end{align}
 
@@
 $\I \omega_{\phi,\pm} = \omega_{\phi,\pm}$ in the basis of the drive
-operators, finding
+operators $D_{LM}$, finding
 \begin{align}

### 2019-09-27 collective_sun/sun_notes.tex
fixed minor error in definition of Dicke states

@@
 magnetic field and a classical driving laser; loosely speaking, one
-should identify $\p{D_{1,0},D_{1,1}}\sim\p{S_\z,S_\x}$, where
-$S_\alpha$ is a collective SU(2) spin-$\alpha$
+can identify $\p{D_{1,0},D_{1,1}}\sim\p{S_\z,S_\x}$, where $S_\alpha$
+is a collective SU(2) spin-$\alpha$
 operator\cite{perlin2019shorttime}.  More concretely, we can embed the
@@
 \begin{align}
-  \ket{\mu}_{\t{single}} \to \ket{\mu+I}_{\t{collective}}
-  \propto S_+ \ket\dn^{\otimes\p{n-1}},
+  \ket{\mu}_{\t{single}} \to \ket{I+\mu}_{\t{collective}}
+  \propto S_+^{I+\mu} \ket\dn^{\otimes\p{n-1}},
 \end{align}
@@
 
-An external driving laser addressing a spin-$1/2$ particle on a
-lattice will induce the Hamiltonian
+An external driving laser addressing a spin-$1/2$ fermion on a lattice
+will induce the Hamiltonian
 \begin{align}
@@
 where $\D_{LM}\equiv\sum_jD_{LM}^{(j)}/2$ is a collective drive
-operator, with e.g.~$\D_{1,1}=S_\x$ in the case of $n=2$.  The SU($n$)
-symmetry of collective interactions implies that the interaction
-Hamiltonian $H_{\t{int}}$ in \eqref{eq:H_int_drive} is invariant under
-this gauge transformation, but the single-particle Hamiltonian in
-\eqref{eq:H_lat} transforms as
+operator.  The SU($n$) symmetry of collective interactions implies
+that the interaction Hamiltonian $H_{\t{int}}$ in
+\eqref{eq:H_int_drive} is invariant under this gauge transformation,
+but the single-particle Hamiltonian in \eqref{eq:H_lat} transforms as
 \begin{align}

### 2019-10-01 collective_sun/sun_notes.tex
fixed minor error in definition of drive operators
add ideas for future directions to take the current work
changed some notation and added (i) transition operator algebra structure constants (ii) Hamiltonian for a Raman drive

@@
 
+\usepackage[inline]{enumitem} % in-line lists and \setlist{} (below)
+\setlist{nolistsep} % more compact spacing between environments
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \section{Collective SU($n$)-symmetric interactions}
+\label{sec:int}
 
@@
 \begin{align}
-  H_{\t{int}} = \f12 u \sum_{p,q,\mu,\nu}
+  H_{\t{int}} = \f{u}{2} \sum_{p,q,\mu,\nu}
   \p{c_{p\mu}^\dag c_{q\nu}^\dag c_{q\nu} c_{p\mu}
@@
 \begin{align}
-  H_{\t{int}} = \f12 u \sum_{p,q,\mu,\nu}
+  H_{\t{int}} = \f{u}{2} \sum_{p,q,\mu,\nu}
   \p{s_{\mu\mu}^{(p)} s_{\nu\nu}^{(q)} - s_{\mu\nu}^{(p)} s_{\nu\mu}^{(q)}}
-  = \f12 u\p{N^2 - \sum_{p,q,\mu,\nu} s_{\mu\nu}^{(p)} s_{\nu\mu}^{(q)}},
+  = \f{u}{2}\p{N^2 - \sum_{p,q,\mu,\nu} s_{\mu\nu}^{(p)} s_{\nu\mu}^{(q)}},
 \end{align}
@@
   H_{\t{int}}
-  = - \f12 u \sum_{p,q,\mu,\nu} s_{\mu\nu}^{(p)} s_{\nu\mu}^{(q)}.
+  = - \f{u}{2} \sum_{p,q,\mu,\nu} s_{\mu\nu}^{(p)} s_{\nu\mu}^{(q)}.
 \end{align}
@@
   H_{\t{int}}
-  = - \f12 u \sum_{p,q} {\v s^{(p)}}^\dag \v s^{(q)}
-  = - \f12 u \v\S^\dag \c \v\S,
+  = - \f{u}{2} \sum_{p,q} {\v s^{(p)}}^\dag \v s^{(q)}
+  = - \f{u}{2} \v\S^\dag \c \v\S,
   &&
@@
 \subsection{Generalized Gell-Mann (GGM) operators}
+\label{sec:ggm_ops}
 
@@
   H_{\t{int}}
-  = -\f14 u \sum_{p,q} \v\lambda^{(p)} \c \v\lambda^{(q)}
-  = -u \v\Lambda \c \v\Lambda,
+  = -\f{u}{4} \sum_{p,q} \v\lambda^{(p)} \c \v\lambda^{(q)}
+  = -\f{u}{4} \v\Lambda \c \v\Lambda,
   &&
-  \v\Lambda \equiv \f12 \sum_p \v\lambda^{(p)}.
+  \v\Lambda \equiv \sum_p \v\lambda^{(p)}.
   \label{eq:H_int_GGM}
@@
 \subsection{Transition operators}
+\label{sec:trans_ops}
 
@@
 \begin{align}
-  T_{LM} \equiv \sqrt{\f{2L+1}{2I+1}} \sum_{\mu,\nu}
-  \bk{I\mu;I\nu,LM} \op{\mu}{\nu},
+  T_{LM}
+  \equiv \sqrt{\f{2L+1}{2I+1}} \sum_{\mu,\nu}
+  \bk{I\mu,LM;I\nu} \op{\nu}{\mu},
   \label{eq:trans_ops}
@@
 \begin{align}
-  H_{\t{int}} = -\f12 u \sum_{p,q} {\v T^{(p)}}^\dag \c \v T^{(q)}
-  = -\f12 u \v\T^\dag \c \v\T,
+  H_{\t{int}} = -\f{u}{2} \sum_{p,q} {\v T^{(p)}}^\dag \c \v T^{(q)}
+  = -\f{u}{2} \v\T^\dag \c \v\T,
   &&
@@
 \subsection{Drive operators}
+\label{sec:drive_ops}
 
@@
   &&
-  D_{LM} \stackrel{M>0}{\equiv} -\p{T_{L\Delta} + T_{L\Delta}^\dag},
+  D_{LM} \stackrel{M>0}{\equiv} -\p{T_{L\abs{M}} + T_{L\abs{M}}^\dag},
   &&
-  D_{LM} \stackrel{M<0}{\equiv} i\p{T_{L\Delta} - T_{L\Delta}^\dag},
+  D_{LM} \stackrel{M<0}{\equiv} i\p{T_{L\abs{M}} - T_{L\abs{M}}^\dag},
   \label{eq:drive_ops}
@@
 \begin{align}
-  H_{\t{int}} = -\f14 u \sum_{p,q} \v D^{(p)}\c\v D^{(q)}
-  = -u \v\D \c \v\D,
+  H_{\t{int}} = -\f{u}{4} \sum_{p,q} \v D^{(p)}\c\v D^{(q)}
+  = -\f{u}{4} \v\D \c \v\D,
   &&
-  \v\D \equiv \f12 \sum_p \v D^{(p)},
+  \v\D \equiv \sum_p \v D^{(p)},
   \label{eq:H_int_drive}
@@
 \section{External drive and spin-orbit coupling}
+\label{sec:drive_SOC}
 
@@
   \left. H_{\t{drive}}^{(\phi)} \right|_{n=2}
-  = \f12 \Omega \sum_j
+  = \Omega \sum_j
   \p{e^{i\phi_j} c_{j,\up}^\dag c_{j,\dn}
@@
   H_{\t{drive}}^{(\phi)}
-  = -\f12 \Omega \sum_j \p{e^{i\phi_j} T_{1,1}^{(j)}
+  = -\Omega \sum_j \p{e^{i\phi_j} T_{1,1}^{(j)}
     + e^{-i\phi_j} {T_{1,1}^{(j)}}^\dag}.
@@
   H_{\t{drive}}
-  = -\f12 \Omega \sum_j\p{T_{1,1}^{(j)} + {T_{1,1}^{(j)}}^\dag}
-  = \f12 \Omega \sum_j D_{1,1}
+  = -\Omega \sum_j\p{T_{1,1}^{(j)} + {T_{1,1}^{(j)}}^\dag}
+  = \Omega \sum_j D_{1,1}
   = \Omega \D_{1,1},
 \end{align}
-where $\D_{LM}\equiv\sum_jD_{LM}^{(j)}/2$ is a collective drive
+where $\D_{LM}\equiv\sum_jD_{LM}^{(j)}$ is a collective drive
 operator.  The SU($n$) symmetry of collective interactions implies
@@
   B_L\p{q,\phi} = \p{-1}^L f_L\p{qa} \sqrt{\f{L+1/2}{n}}
-  \sum_\mu \bk{I\mu;I\mu,L,0} f_L\p{\mu\phi},
+  \sum_\mu \bk{I\mu,L,0;I\mu} f_L\p{\mu\phi},
   &&
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Classifying eigenstates}
+\section{Raman drive}
+\label{sec:raman}
 
-[not yet written]
+Rather than coupling states with a single driving laser directly, as
+in Section \ref{sec:drive_SOC}, we now consider coupling through an
+off-resonant Raman transition.  Specifically, we consider a
+three-laser drive of the form
+\begin{align}
+  H_{\t{Raman,bare}}^{(\phi)}
+  = \sum_j \sum_{m\in\set{-1,0,1}} \Omega_m T_{1,m}^{(j)}
+  \otimes \sigma_+^{(j)} e^{i\phi_{jm}-i\Delta t} + \t{h.c.},
+\end{align}
+where $j$ indexes a lattice site, $\Omega_m$ is a driving
+strength, $\sigma_+$ is a spin-raising (Pauli) operator, $\phi_{jm}$
+is a phase, and $\Delta$ is a detuning of the lasers from the energy
+of an excited state.  If $\Delta\gg\Omega_m$, then at second
+order in perturbation theory we can find that the effective action of
+$H_{\t{Raman,bare}}^{(\phi)}$ is
+\begin{align}
+  H_{\t{Raman}}^{(\phi)}
+  = \sum_{\tau\in\set{0,1,2}} H_{\t{Raman},\tau}^{(\phi)}
+\end{align}
+with
+\begin{align}
+  H_{\t{Raman},0}^{(\phi)}
+  &\equiv -\sum_j\sum_m \f{\Omega_m^2}{\Delta}
+  {T_{1,m}^{(j)}}^\dag T_{1,m}^{(j)}, & \label{eq:raman_0} \\
+  H_{\t{Raman},1}^{(\phi)}
+  &\equiv -\sum_j\sum_{s\in\set{\pm1}}
+  \f{\Omega_0\Omega_s}{\Delta}
+  e^{i\p{\phi_{js}-\phi_{j,0}}}{T_{1,0}^{(j)}}^\dag T_{1,s}^{(j)} + \t{h.c.},
+  \\
+  H_{\t{Raman},2}^{(\phi)}
+  &\equiv -\f{\Omega_1\Omega_{-1}}{\Delta}
+  \sum_j\sum_{s\in\set{\pm1}}
+  e^{i\p{\phi_{js}-\phi_{j,-s}}} {T_{1,-s}^{(j)}}^\dag T_{1,s}^{(j)}.
+  \label{eq:raman_2}
+\end{align}
+In terms of the drive operators, we can expand
+\begin{align}
+  H_{\t{Raman}}^{(\phi)}
+  = \sum_j \sum_{\substack{L\in\set{0,1,2}\\\abs{m}\le L}}
+  \Omega_{LM}^{(\phi,j)} D_{LM}^{(j)},
+\end{align}
+where the coefficients $\Omega_{LM}^{(\phi,j)}$ are provided in
+Appendix \ref{sec:raman_coeff}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Future directions}
+
+\begin{itemize}
+\item Implications of SOC
+  \begin{itemize}
+  \item Perturbative regime: $\phi\ll1$
+  \item Special angles $\phi$
+  \end{itemize}
+\item Squeezing
+  \begin{itemize}
+  \item Collection of $n$-level spins
+  \item Generalizations of squeezing beyond SU(2)
+  \end{itemize}
+\item Consider a different type of drive
+\item Consider different interactions
+  \begin{itemize}
+  \item Include electronic states
+  \item Super-exchange regime
+  \item Non-uniform (``disordered'') interactions
+  \end{itemize}
+\end{itemize}
 
@@
 \end{align}
-To maintain generality, we choose an arbitrary basis $\set{X_j}$
-satisfying the orthonormality condition
+To maintain generality, we choose an arbitrary basis
+$X\equiv\set{X_j}$ satisfying the orthonormality condition
 \begin{align}
@@
 \end{align}
-Choosing a different basis $\set{Y_j}$ for $\B_n$ with corresponding
-norm $\N_Y$, we can expand
+Choosing a different basis $Y\equiv\set{Y_j}$ for $\B_n$ with
+corresponding norm $\N_Y$, we can expand
 \begin{align}
@@
 $\I \omega_{\phi,\pm} = \omega_{\phi,\pm}$ in the basis of the drive
-operators $D_{LM}$, finding
+operators $D_{LM}$, finding (see Appendix \ref{sec:changing_bases})
 \begin{align}
@@
   &= \p{-1}^L f_L\p{qa} \sqrt{\f{L+1/2}{n}}
-  \sum_\mu \bk{I\mu;I\mu,L,0} f_L\p{\mu\phi},
+  \sum_\mu \bk{I\mu,L,0;I\mu} f_L\p{\mu\phi},
 \end{align}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Transition operator product expansion}
+\label{sec:trans_prod}
+
+Here we compute, for an $n$-level spin with total spin
+$I\equiv\p{n-1}/2$, the structure constants of the transition operator
+algebra, expanding
+\begin{align}
+  T_{\ell_2 m_2} T_{\ell_1 m_1}
+  = \sum_{L,M} T_{LM} \obk{T_{LM} | T_{\ell_2 m_2} T_{\ell_1 m_1}},
+\end{align}
+where the transition operators are defined by
+\begin{align}
+  T_{LM} \equiv \sqrt{\f{2L+1}{2I+1}}
+  \sum_{\mu,\nu} \bk{I\mu,LM;I\nu} \op{\nu}{\mu},
+\end{align}
+with Clebsch-Gordan coefficients $\bk{I\mu,LM;I\nu}$, and the inner
+product is
+\begin{align}
+  \obk{T_{LM} | T_{\ell_2 m_2} T_{\ell_1 m_1}}
+  = \tr\p{T_{LM}^\dag T_{\ell_2 m_2} T_{\ell_1 m_1}}
+  = \p{-1}^M \tr\p{T_{L,-M} T_{\ell_2 m_2} T_{\ell_1 m_1}}.
+\end{align}
+The Clebsch-Gordan coefficients can be written in terms of Wigner
+3-$j$ symbols as
+\begin{align}
+  \bk{\ell_1 m_1, \ell_2 m_2;\ell_3 m_3}
+  = \p{-1}^{-\ell_1+\ell_2-m_3} \sqrt{\p{2\ell_3+1}}
+  \begin{pmatrix}
+    \ell_1 & \ell_2 & \ell_3 \\
+    m_1 & m_2 & -m_3
+  \end{pmatrix}.
+\end{align}
+Substituting transition operators and replacing Clebsch-Gordan
+coefficients by Wigner 3-$j$ symbols, we find that
+\begin{multline}
+  \obk{T_{LM} | T_{\ell_2 m_2} T_{\ell_1 m_1}}
+  = \delta_{m_1+m_2,M}
+  \p{-1}^M \sqrt{\p{2\ell_1+1}\p{2\ell_2+1}\p{2L+1}} \\
+  \times \sum_{\mu,\nu,\rho} \p{-1}^{-3I+\ell_1+\ell_2+L-\mu-\nu-\rho}
+  \begin{pmatrix}
+    I & \ell_1 & I \\
+    \mu & m_1 & -\nu
+  \end{pmatrix}
+  \begin{pmatrix}
+    I & \ell_2 & I \\
+    \nu & m_2 & -\rho
+  \end{pmatrix}
+  \begin{pmatrix}
+    I & L & I \\
+    \rho & -M & -\mu
+  \end{pmatrix},
+\end{multline}
+where we can evaluate the sum and introduce Wigner 6-$j$ symbols to
+get
+\begin{multline}
+  \obk{T_{LM} | T_{\ell_2 m_2} T_{\ell_1 m_1}} \\
+  = \delta_{m_1+m_2,M} \p{-1}^{2I+M}
+  \sqrt{\p{2\ell_1+1}\p{2\ell_2+1}\p{2L+1}}
+  \begin{pmatrix}
+    \ell_1 & \ell_2 & L \\
+    m_1 & m_2 & -M
+  \end{pmatrix}
+  \begin{Bmatrix}
+    \ell_1 & \ell_2 & L \\
+    J & J & J
+  \end{Bmatrix} \\
+  = \delta_{m_1+m_2,M} \p{-1}^{2I+\ell_1-\ell_2}
+  \sqrt{\p{2\ell_1+1}\p{2\ell_2+1}}
+  \bk{\ell_1 m_2, \ell_2 m_2; LM}
+  \begin{Bmatrix}
+    \ell_1 & \ell_2 & L \\
+    J & J & J
+  \end{Bmatrix}.
+\end{multline}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Raman drive coefficients}
+\label{sec:raman_coeff}
+
+Here we provide the coefficients $\Omega_{LM}^{(\phi,j)}$ of the
+effective Raman driving Hamiltonian
+\begin{align}
+  H_{\t{Raman}}^{(\phi)}
+  = \sum_j \sum_{\substack{L\in\set{0,1,2}\\\abs{m}\le L}}
+  \Omega_{LM}^{(\phi,j)} D_{LM}^{(j)}
+\end{align}
+considered in Section \ref{sec:raman}.  These coefficients are
+computed by evaluating products of transition operators in
+\eqref{eq:raman_0}--\eqref{eq:raman_2}, using both the structure
+constants of the transition operator algebra (derived in Appendix
+\ref{sec:trans_prod}) and the fact that
+$T_{LM}^\dag = \p{-1}^M T_{L,-M}$.  For clarity, if any coefficient
+$\Omega_{LM}\p{\phi}$ is independent of $\phi$, we drop the explicit
+dependence on $\phi$, i.e.~taking
+$\Omega_{LM}^{(\phi,j)}\to\Omega_{LM}$.
+\begin{align}
+  \Omega_{0,0} \equiv -\f1{\sqrt{2n}} \sum_m\f{\Omega_m^2}{\Delta},
+  &&
+  \Omega_{1,0} \equiv \sqrt{\f{3}{2n\p{n^2-1}}}
+  \sum_m m\f{\Omega_m^2}{\Delta},
+\end{align}
+\begin{align}
+  \Omega_{2,0} \equiv -\sqrt{\f{\p{n^2-4}}{10n\p{n^2-1}}}
+  \sum_m \p{-1}^m \p{1+\delta_{m,0}} \f{\Omega_m^2}{\Delta},
+\end{align}
+\begin{align}
+  \Omega_{L,1}^{(\phi,j)}
+  \equiv \sum_{s\in\set{\pm1}} \omega_{Ls} \cos\p{\phi_{js}-\phi_{j,0}},
+  &&
+  \Omega_{L,-1}^{(\phi,j)}
+  \equiv -\sum_{s\in\set{\pm1}}
+  s~ \omega_{Ls} \sin\p{\phi_{js}-\phi_{j,0}},
+\end{align}
+\begin{align}
+  \omega_{1,s} \equiv \f{\sqrt3}{\sqrt{n\p{n^2-1}}}
+  \f{\Omega_0 \Omega_s}{\Delta},
+  &&
+  \omega_{2,s} \equiv s~ \f{\sqrt{3/5}}{n\p{n^2-1}}
+  \sqrt{\f{\p{n+2}!}{\p{n-3}!}}~ \f{\Omega_0\Omega_s}{\Delta},
+\end{align}
+\begin{align}
+  \Omega_{2,2}^{(\phi,j)}
+  \equiv \omega_{2,2} \cos\p{\phi_{j,1}-\phi_{j,-1}},
+  &&
+  \Omega_{2,-2}^{(\phi,j)}
+  \equiv -\omega_{2,2} \sin\p{\phi_{j,1}-\phi_{j,-1}},
+\end{align}
+\begin{align}
+  \omega_{2,2} \equiv -\f{\sqrt{6/5}}{n\p{n^2-1}}
+  \sqrt{\f{\p{n+2}!}{\p{n-3}!}} \f{\Omega_1\Omega_{-1}}{\Delta}.
+\end{align}
+
 \bibliography{sun_notes.bib}

### 2019-10-02 collective_sun/sun_notes.tex
change sign conventions on drive laser phases

@@
   = \Omega \sum_j
-  \p{e^{i\phi_j} c_{j,\up}^\dag c_{j,\dn}
-    + e^{-i\phi_j} c_{j,\dn}^\dag c_{j,\up}},
+  \p{e^{-i\phi_j} c_{j,\up}^\dag c_{j,\dn}
+    + e^{i\phi_j} c_{j,\dn}^\dag c_{j,\up}},
   \label{eq:H_drive_2}
@@
   H_{\t{drive}}^{(\phi)}
-  = -\Omega \sum_j \p{e^{i\phi_j} T_{1,1}^{(j)}
-    + e^{-i\phi_j} {T_{1,1}^{(j)}}^\dag}.
+  = -\Omega \sum_j \p{e^{-i\phi_j} T_{1,1}^{(j)}
+    + e^{i\phi_j} {T_{1,1}^{(j)}}^\dag}.
   \label{eq:H_drive}
@@
 We can make this drive spatially homogenous via the gauge
-transformation $c_{j\mu} \to e^{i\mu\phi_j} c_{j\mu}$, after which the
+transformation $c_{j\mu} \to e^{-i\mu\phi_j} c_{j\mu}$, after which the
 drive becomes
@@
   \to H_{\t{lat}}^{(\phi)}
-  = -J \sum_{q,\mu} \cos\p{qa+\mu\phi} c_{q\mu}^\dag c_{q\mu},
+  = -J \sum_{q,\mu} \cos\p{qa-\mu\phi} c_{q\mu}^\dag c_{q\mu},
   \label{eq:H_lat_SOC}
@@
 \begin{align}
-  B_L\p{q,\phi} = \p{-1}^L f_L\p{qa} \sqrt{\f{L+1/2}{n}}
+  B_L\p{q,\phi} = f_L\p{qa} \sqrt{\f{L+1/2}{n}}
   \sum_\mu \bk{I\mu,L,0;I\mu} f_L\p{\mu\phi},
@@
 in Section \ref{sec:drive_SOC}, we now consider coupling through an
-off-resonant Raman transition.  Specifically, we consider a
-three-laser drive of the form
+off-resonant Raman transition to a highly excited state.
+Specifically, we consider a three-laser drive of the form
 \begin{align}
   H_{\t{Raman,bare}}^{(\phi)}
-  = \sum_j \sum_{m\in\set{-1,0,1}} \Omega_m T_{1,m}^{(j)}
-  \otimes \sigma_+^{(j)} e^{i\phi_{jm}-i\Delta t} + \t{h.c.},
+  = \sum_j \sum_{m\in\set{-1,0,1}} \p{\Omega_m e^{-i\phi_{jm}}
+    T_{1,m}^{(j)} \otimes \op{\up}{\dn}^{(j)} + \t{h.c.}}
+  + \f{\Delta}{2} \sum_j \1^{(j)}
+  \otimes \p{\op{\up}^{(j)} - \op{\dn}^{(j)}},
+  \label{eq:raman_bare}
 \end{align}
-where $j$ indexes a lattice site, $\Omega_m$ is a driving
-strength, $\sigma_+$ is a spin-raising (Pauli) operator, $\phi_{jm}$
-is a phase, and $\Delta$ is a detuning of the lasers from the energy
-of an excited state.  If $\Delta\gg\Omega_m$, then at second
+where $j$ indexes a lattice site, $\Omega_m$ is a driving strength;
+$\phi_{jm}$ is a site- and drive-dependent phase; $\Delta$ is a
+detuning of the lasers from the energy of an excited state; and $\1$
+is the identity operator.  If $\Delta\gg\Omega_m$, then at second
 order in perturbation theory we can find that the effective action of
-$H_{\t{Raman,bare}}^{(\phi)}$ is
+$H_{\t{Raman,bare}}^{(\phi)}$ within the $\ket{\dn}$-state manifold is
 \begin{align}
@@
   &\equiv -\sum_j\sum_{s\in\set{\pm1}}
-  \f{\Omega_0\Omega_s}{\Delta}
-  e^{i\p{\phi_{js}-\phi_{j,0}}}{T_{1,0}^{(j)}}^\dag T_{1,s}^{(j)} + \t{h.c.},
+  \f{\Omega_0\Omega_s}{\Delta} e^{-i\phi_{js}}
+  {T_{1,0}^{(j)}}^\dag T_{1,s}^{(j)} + \t{h.c.},
   \\
@@
   &\equiv -\f{\Omega_1\Omega_{-1}}{\Delta}
-  \sum_j\sum_{s\in\set{\pm1}}
-  e^{i\p{\phi_{js}-\phi_{j,-s}}} {T_{1,-s}^{(j)}}^\dag T_{1,s}^{(j)}.
+  \sum_j\sum_{s\in\set{\pm1}} e^{-i\p{\phi_{js}-\phi_{j,-s}}}
+  {T_{1,-s}^{(j)}}^\dag T_{1,s}^{(j)},
   \label{eq:raman_2}
 \end{align}
-In terms of the drive operators, we can expand
+where by choice of gauge we set $\phi_{j,0}=0$, or equivalently take
+$\phi_{jm}\to\phi_{jm}-\phi_{j,0}$.  In terms of the drive operators,
+we can expand
 \begin{align}
   H_{\t{Raman}}^{(\phi)}
-  = \sum_j \sum_{\substack{L\in\set{0,1,2}\\\abs{m}\le L}}
+  = \sum_j \sum_{\substack{L\le2\\\abs{m}\le L}}
   \Omega_{LM}^{(\phi,j)} D_{LM}^{(j)},
+  \label{eq:raman_drive}
 \end{align}
@@
   H_{\t{lat}}^{(\phi)}
-  = -J \sum_{q,\mu} \cos\p{qa+\mu\phi} s_{\mu\mu}^{(q)}
+  = -J \sum_{q,\mu} \cos\p{qa-\mu\phi} s_{\mu\mu}^{(q)}
   \label{eq:H_lat_spin}
@@
   \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
-  = \cos\p{qa} \omega_{\phi,+} - \sin\p{qa} \omega_{\phi,-},
+  = \cos\p{qa} \omega_{\phi,+} + \sin\p{qa} \omega_{\phi,-},
 \end{align}
@@
   = \cos\p{qa} \f12 \sum_{L,M} \obk{\omega_{\phi,+}|D_{LM}} D_{LM}
-  - \sin\p{qa} \f12 \sum_{L,M} \obk{\omega_{\phi,-}|D_{LM}} D_{LM}.
+  + \sin\p{qa} \f12 \sum_{L,M} \obk{\omega_{\phi,-}|D_{LM}} D_{LM}.
 \end{align}
@@
   \obk{\omega_{\phi,+}|D_{L,0}} D_{L,0}
-  - \sin\p{qa} \f12 \sum_{L\,\t{odd}}
+  + \sin\p{qa} \f12 \sum_{L\,\t{odd}}
   \obk{\omega_{\phi,-}|D_{L,0}} D_{L,0},
@@
   B_L\p{q,\phi}
-  &\equiv \p{-1}^L f_L\p{qa} \f12
-  \sum_\mu f_L\p{\mu\phi} \obk{s_{\mu\mu}|D_{L,0}} \\
-  &= \p{-1}^L f_L\p{qa} \sqrt{\f{L+1/2}{n}}
+  \equiv f_L\p{qa} \f12
+  \sum_\mu f_L\p{\mu\phi} \obk{s_{\mu\mu}|D_{L,0}}
+  = f_L\p{qa} \sqrt{\f{L+1/2}{n}}
   \sum_\mu \bk{I\mu,L,0;I\mu} f_L\p{\mu\phi},
@@
   \Omega_{L,1}^{(\phi,j)}
-  \equiv \sum_{s\in\set{\pm1}} \omega_{Ls} \cos\p{\phi_{js}-\phi_{j,0}},
+  \equiv \sum_{s\in\set{\pm1}} \omega_{Ls} \cos\phi_{js},
   &&
   \Omega_{L,-1}^{(\phi,j)}
-  \equiv -\sum_{s\in\set{\pm1}}
-  s~ \omega_{Ls} \sin\p{\phi_{js}-\phi_{j,0}},
+  \equiv \sum_{s\in\set{\pm1}} s~ \omega_{Ls} \sin\phi_{js},
 \end{align}
@@
   \Omega_{2,2}^{(\phi,j)}
-  \equiv \omega_{2,2} \cos\p{\phi_{j,1}-\phi_{j,-1}},
+  \equiv \omega_{2,2} \cos\p{\phi_{j+}-\phi_{j-}},
   &&
   \Omega_{2,-2}^{(\phi,j)}
-  \equiv -\omega_{2,2} \sin\p{\phi_{j,1}-\phi_{j,-1}},
+  \equiv \omega_{2,2} \sin\p{\phi_{j+}-\phi_{j-}},
 \end{align}
@@
 \end{align}
+Here $\Omega_m$, $\phi_{jm}$, and $\Delta$ are respectively the drive
+strengths, phases, and detuning that appear in the ``bare'' Raman
+drive in \eqref{eq:raman_bare}.
 

### 2019-10-03 collective_sun/sun_notes.tex
added a bit of discussion for transition and drive operators added coefficients and driving Hamiltonian for a phase-matched Raman drive

@@
 $n$-level nuclear spin onto a quantization axis; and
-$\bk{I\mu;I\nu,LM}$ is a Clebsch-Gordan coefficient.  Similarly to the
-GGM operators in \eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag}, the
+$\bk{I\mu;I\nu,LM}$ is a Clebsch-Gordan coefficient.  The transition
+operator $T_{LM}$ is proportional to the transition induced on a
+nuclear spin by the absorption of a spin-$L$ particle with spin
+projection $M$ onto a quantization axis.  Similarly to the GGM
+operators in \eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag}, the
 transition operators in \eqref{eq:trans_ops} provide a complete basis
@@
 \end{align}
-for integer $M$ with $\abs{M}\le L$.  Similarly to the GGM operators
-in \eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag}, the drive
-operators are equal to the Pauli operators in the case of $n=2$, and
-satisfy the orthonormality condition
+for integer $M$ with $\abs{M}\le L$.  The drive operator $D_{LM}$ is
+proportional to the Hamiltonian induced on a nuclear spin by an
+external classical field of spin-$L$ particles with spin projection
+$M$ onto a quantization axis.  Similarly to the GGM operators in
+\eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag}, the drive operators
+are equal to the Pauli operators in the case of $n=2$, and satisfy the
+orthonormality condition
 \begin{align}
@@
 \end{align}
-where $a,b$ index any drive operator in \eqref{eq:drive_ops}.  In
-terms of the drive operators,
+which implies
 \begin{align}
@@
 
-The drive operators have the convenient feature that $D_{1,0}$ and
-$D_{1,1}$ are (up to a gauge transformation) respectively proportional
-to the Hamiltonians induced on an $n$-level spin by an external
-magnetic field and a classical driving laser; loosely speaking, one
-can identify $\p{D_{1,0},D_{1,1}}\sim\p{S_\z,S_\x}$, where $S_\alpha$
-is a collective SU(2) spin-$\alpha$
+The drive operators $D_{1,0}$ and $D_{1,\pm1}$ are proportional to the
+Hamiltonians induced on an $n$-level spin by an external magnetic
+field and a classical spin-polarized driving laser; loosely speaking,
+one can identify $\p{D_{1,0},D_{1,1},D_{1,-1}}\sim\p{S_\z,S_\x,S_\y}$,
+where $S_\alpha$ is a collective SU(2) spin-$\alpha$
 operator\cite{perlin2019shorttime}.  More concretely, we can embed the
@@
 
-An external driving laser addressing a spin-$1/2$ fermion on a lattice
-will induce the Hamiltonian
+An driving laser addressing spin-$1/2$ fermions on a 1-D lattice will
+induce the Hamiltonian
 \begin{align}
@@
 \end{align}
-where $\Omega$ is a drive strength, $j$ indexes a lattice site, and
-$\phi_j$ is a site-dependent phase.  The multi-level analogue of
-\eqref{eq:H_drive_2} is
+where $\Omega$ is a (real) driving amplitude, $j$ indexes a lattice
+site, and $\phi_j$ is a site-dependent phase.  The multi-level
+analogue of \eqref{eq:H_drive_2} is
 \begin{align}
@@
   = -\Omega \sum_j \p{e^{-i\phi_j} T_{1,1}^{(j)}
-    + e^{i\phi_j} {T_{1,1}^{(j)}}^\dag}.
+    + e^{i\phi_j} {T_{1,1}^{(j)}}^\dag}
+  = \Omega \D_{1,1}^{(\phi)},
   \label{eq:H_drive}
 \end{align}
-We can make this drive spatially homogenous via the gauge
-transformation $c_{j\mu} \to e^{-i\mu\phi_j} c_{j\mu}$, after which the
-drive becomes
+where
 \begin{align}
-  H_{\t{drive}}
-  = -\Omega \sum_j\p{T_{1,1}^{(j)} + {T_{1,1}^{(j)}}^\dag}
-  = \Omega \sum_j D_{1,1}
-  = \Omega \D_{1,1},
+  \D_{LM}^{(\phi)}
+  \equiv \sum_j \sp{\cos\p{M\phi_j} D_{LM}^{(j)}
+    + \sin\p{M\phi_j} D_{L,-M}^{(j)}},
+  \label{eq:drive_rot}
 \end{align}
-where $\D_{LM}\equiv\sum_jD_{LM}^{(j)}$ is a collective drive
-operator.  The SU($n$) symmetry of collective interactions implies
-that the interaction Hamiltonian $H_{\t{int}}$ in
-\eqref{eq:H_int_drive} is invariant under this gauge transformation,
-but the single-particle Hamiltonian in \eqref{eq:H_lat} transforms as
+is an inhomogeneously rotated collective drive operator.  We can make
+this drive spatially homogenous via the gauge transformation
+$c_{j\mu}^\dag \to e^{i\mu\phi_j} c_{j\mu}^\dag$, after which the
+drive becomes simply
+\begin{align}
+  H_{\t{drive}} = \Omega \D_{1,1}.
+\end{align}
+The SU($n$) symmetry of collective interactions implies that the
+interaction Hamiltonian $H_{\t{int}}$ in \eqref{eq:H_int_drive} is
+invariant under this gauge transformation, but the single-particle
+Hamiltonian in \eqref{eq:H_lat} transforms as
 \begin{align}
@@
 \section{Raman drive}
-\label{sec:raman}
+\label{sec:drive_raman}
 
-Rather than coupling states with a single driving laser directly, as
-in Section \ref{sec:drive_SOC}, we now consider coupling through an
-off-resonant Raman transition to a highly excited state.
-Specifically, we consider a three-laser drive of the form
+Rather than coupling the states of a multi-level system directly with
+a single driving laser, as in Section \ref{sec:drive_SOC}, we now
+consider coupling through an off-resonant Raman transition to a highly
+excited state.  Specifically, we consider a three-laser drive of the
+form
 \begin{align}
   H_{\t{Raman,bare}}^{(\phi)}
-  = \sum_j \sum_{m\in\set{-1,0,1}} \p{\Omega_m e^{-i\phi_{jm}}
-    T_{1,m}^{(j)} \otimes \op{\up}{\dn}^{(j)} + \t{h.c.}}
-  + \f{\Delta}{2} \sum_j \1^{(j)}
-  \otimes \p{\op{\up}^{(j)} - \op{\dn}^{(j)}},
+  = \sum_j \sum_{m\in\set{1,0,-1}} \p{\Omega_m e^{-i\phi_{jm}}
+    T_{1,m}^{(j)} \otimes \op{\up}{\dn} + \t{h.c.}}
+  + \f{\Delta}{2} \1 \otimes \p{\op{\up} - \op{\dn}},
   \label{eq:raman_bare}
 \end{align}
-where $j$ indexes a lattice site, $\Omega_m$ is a driving strength;
-$\phi_{jm}$ is a site- and drive-dependent phase; $\Delta$ is a
-detuning of the lasers from the energy of an excited state; and $\1$
-is the identity operator.  If $\Delta\gg\Omega_m$, then at second
-order in perturbation theory we can find that the effective action of
+where $j$ indexes a lattice site, $\Omega_m$ is a (real) driving
+amplitude; $\phi_{jm}$ is a site- and drive-dependent phase;
+$\op{r}{s}$ for $r,s\in\set{\up,\dn}$ is a pseudo-spin operator for
+the auxiliary degree of freedom off-resonantly addressed by the
+lasers; $\Delta$ is a detuning of the lasers from the energy of an
+excited state; and $\1$ is the identity operator.  If the detuning is
+large, $\Delta\gg\Omega_m$, then at second order in perturbation
+theory we can find that the effective action of
 $H_{\t{Raman,bare}}^{(\phi)}$ within the $\ket{\dn}$-state manifold is
@@
   H_{\t{Raman},2}^{(\phi)}
-  &\equiv -\f{\Omega_1\Omega_{-1}}{\Delta}
-  \sum_j\sum_{s\in\set{\pm1}} e^{-i\p{\phi_{js}-\phi_{j,-s}}}
+  &\equiv -\f{\Omega_+\Omega_-}{\Delta}
+  \sum_j\sum_{s\in\set{\pm1}} e^{-is\p{\phi_{j+}-\phi_{j-}}}
   {T_{1,-s}^{(j)}}^\dag T_{1,s}^{(j)},
@@
 \end{align}
-where by choice of gauge we set $\phi_{j,0}=0$, or equivalently take
-$\phi_{jm}\to\phi_{jm}-\phi_{j,0}$.  In terms of the drive operators,
+where by choice of gauge we set $\phi_{j,0}=0$, or equivalently
+re-define $\phi_{jm}-\phi_{j,0}\to\phi_{jm}$, and we identify the
+subscripts $\pm\leftrightarrow\pm1$.  In terms of the drive operators,
 we can expand
@@
   H_{\t{Raman}}^{(\phi)}
-  = \sum_j \sum_{\substack{L\le2\\\abs{m}\le L}}
+  = \sum_j \sum_{\substack{L\le2\\\abs{M}\le L}}
   \Omega_{LM}^{(\phi,j)} D_{LM}^{(j)},
@@
 where the coefficients $\Omega_{LM}^{(\phi,j)}$ are provided in
-Appendix \ref{sec:raman_coeff}.
+Appendix \ref{sec:drive_raman_coeff}.  If the driving lasers are
+phase-matched with $\phi_{js}\equiv s\phi_j$, then the drive further
+simplifies to
+\begin{align}
+  H_{\t{Raman}}^{(\phi)}
+  = \sum_{\substack{L\le2\\0\le M\le L}} \Omega_{LM} \D_{LM}^{(\phi)}
+  \label{eq:drive_raman_matched}
+\end{align}
+with rotated collective drive operators $\D_{LM}^{(\phi)}$ defined in
+\eqref{eq:drive_rot}, and uniform coefficients $\Omega_{LM}$ that have
+the scaling
+\begin{align}
+  \Omega_{0,0} &\propto \Omega_+^2 + \Omega_-^2 + \Omega_0^2,
+  &
+  \Omega_{1,0} &\propto \Omega_+^2 - \Omega_-^2,
+  &
+  \Omega_{1,1} &\propto \Omega_0 \p{\Omega_+ + \Omega_-}, \\
+  \Omega_{2,0} &\propto \Omega_+^2 + \Omega_-^2 - 2\Omega_0^2,
+  &
+  \Omega_{2,1} &\propto \Omega_0 \p{\Omega_+ - \Omega_-},
+  &
+  \Omega_{2,2} &\propto \Omega_+ \Omega_-.
+\end{align}
+Similarly to the case of a direct drive considered in Section
+\ref{sec:drive_SOC}, the Raman drive $H_{\t{Raman}}^{(\phi)}$ in
+\eqref{eq:drive_raman_matched} can be made homogenous through the
+gauge transformation $c_{j\mu}^\dag \to e^{i\mu\phi_j} c_{j\mu}^\dag$,
+under which
+\begin{align}
+  H_{\t{Raman}}^{(\phi)} \to H_{\t{Raman}}
+  \equiv \sum_{\substack{L\le2\\0\le M\le L}} \Omega_{LM} \D_{LM}.
+  \label{eq:drive_raman_homo}
+\end{align}
 
@@
 \begin{align}
-  \omega_{\phi,+} \equiv \sum_\mu \cos\p{\mu\phi} s_{\mu\mu},
+  w_{\phi,+} \equiv \sum_\mu \cos\p{\mu\phi} s_{\mu\mu},
   &&
-  \omega_{\phi,-} \equiv \sum_\mu \sin\p{\mu\phi} s_{\mu\mu},
+  w_{\phi,-} \equiv \sum_\mu \sin\p{\mu\phi} s_{\mu\mu},
 \end{align}
@@
   \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
-  = \cos\p{qa} \omega_{\phi,+} + \sin\p{qa} \omega_{\phi,-},
+  = \cos\p{qa} w_{\phi,+} + \sin\p{qa} w_{\phi,-},
 \end{align}
 and resolve the identity (super-)operator $\I$ with
-$\I \omega_{\phi,\pm} = \omega_{\phi,\pm}$ in the basis of the drive
+$\I w_{\phi,\pm} = w_{\phi,\pm}$ in the basis of the drive
 operators $D_{LM}$, finding (see Appendix \ref{sec:changing_bases})
@@
   \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
-  = \cos\p{qa} \f12 \sum_{L,M} \obk{\omega_{\phi,+}|D_{LM}} D_{LM}
-  + \sin\p{qa} \f12 \sum_{L,M} \obk{\omega_{\phi,-}|D_{LM}} D_{LM}.
+  = \cos\p{qa} \f12 \sum_{L,M} \obk{w_{\phi,+}|D_{LM}} D_{LM}
+  + \sin\p{qa} \f12 \sum_{L,M} \obk{w_{\phi,-}|D_{LM}} D_{LM}.
 \end{align}
@@
 \begin{align}
-  \sum_{L,M} \obk{\omega_{\phi,+}|D_{LM}} D_{LM}
-  = \sum_{L\,\t{even}} \obk{\omega_{\phi,+}|D_{L,0}} D_{L,0},
+  \sum_{L,M} \obk{w_{\phi,+}|D_{LM}} D_{LM}
+  = \sum_{L\,\t{even}} \obk{w_{\phi,+}|D_{L,0}} D_{L,0},
 \end{align}
-and likewise with $\omega_{\phi,-}$ and odd $L$.  In total, we can
+and likewise with $w_{\phi,-}$ and odd $L$.  In total, we can
 write
@@
   = \cos\p{qa} \f12 \sum_{L\,\t{even}}
-  \obk{\omega_{\phi,+}|D_{L,0}} D_{L,0}
+  \obk{w_{\phi,+}|D_{L,0}} D_{L,0}
   + \sin\p{qa} \f12 \sum_{L\,\t{odd}}
-  \obk{\omega_{\phi,-}|D_{L,0}} D_{L,0},
+  \obk{w_{\phi,-}|D_{L,0}} D_{L,0},
 \end{align}
@@
 \section{Raman drive coefficients}
-\label{sec:raman_coeff}
+\label{sec:drive_raman_coeff}
 
@@
   H_{\t{Raman}}^{(\phi)}
-  = \sum_j \sum_{\substack{L\in\set{0,1,2}\\\abs{m}\le L}}
+  = \sum_j \sum_{\substack{L\in\set{0,1,2}\\\abs{M}\le L}}
   \Omega_{LM}^{(\phi,j)} D_{LM}^{(j)}
 \end{align}
-considered in Section \ref{sec:raman}.  These coefficients are
+considered in Section \ref{sec:drive_raman}.  These coefficients are
 computed by evaluating products of transition operators in
@@
 $T_{LM}^\dag = \p{-1}^M T_{L,-M}$.  For clarity, if any coefficient
-$\Omega_{LM}\p{\phi}$ is independent of $\phi$, we drop the explicit
-dependence on $\phi$, i.e.~taking
+$\Omega_{LM}^{(\phi,j)}$ is independent of $\phi$, we drop the
+explicit dependence on $\phi$, i.e.~taking
 $\Omega_{LM}^{(\phi,j)}\to\Omega_{LM}$.
@@
   &&
-  \Omega_{1,0} \equiv \sqrt{\f{3}{2n\p{n^2-1}}}
-  \sum_m m\f{\Omega_m^2}{\Delta},
+  \Omega_{1,0} \equiv \sqrt{\f{3}{2n\p{n^2-1}}}~
+  \f{\Omega_+^2 - \Omega_-}{\Delta},
+  \label{eq:O_01_0}
 \end{align}
 \begin{align}
-  \Omega_{2,0} \equiv -\sqrt{\f{\p{n^2-4}}{10n\p{n^2-1}}}
-  \sum_m \p{-1}^m \p{1+\delta_{m,0}} \f{\Omega_m^2}{\Delta},
+  \Omega_{2,0} \equiv \sqrt{\f{n^2-4}{10n\p{n^2-1}}}~
+  \f{\Omega_+ + \Omega_- - 2\Omega_0}{\Delta},
+  \label{eq:O_2_0}
 \end{align}
@@
   \Omega_{L,-1}^{(\phi,j)}
-  \equiv \sum_{s\in\set{\pm1}} s~ \omega_{Ls} \sin\phi_{js},
-\end{align}
-\begin{align}
-  \omega_{1,s} \equiv \f{\sqrt3}{\sqrt{n\p{n^2-1}}}
-  \f{\Omega_0 \Omega_s}{\Delta},
-  &&
-  \omega_{2,s} \equiv s~ \f{\sqrt{3/5}}{n\p{n^2-1}}
-  \sqrt{\f{\p{n+2}!}{\p{n-3}!}}~ \f{\Omega_0\Omega_s}{\Delta},
+  \equiv \sum_{s\in\set{\pm1}} s \omega_{Ls} \sin\phi_{js},
 \end{align}
@@
 \begin{align}
+  \omega_{1\pm} \equiv \f{\sqrt3}{\sqrt{n\p{n^2-1}}}~
+  \f{\Omega_0 \Omega_\pm}{\Delta},
+  &&
+  \omega_{2\pm} \equiv \pm \f{\sqrt{3/5}}{n\p{n^2-1}}
+  \sqrt{\f{\p{n+2}!}{\p{n-3}!}}~ \f{\Omega_0\Omega_\pm}{\Delta},
+\end{align}
+\begin{align}
   \omega_{2,2} \equiv -\f{\sqrt{6/5}}{n\p{n^2-1}}
-  \sqrt{\f{\p{n+2}!}{\p{n-3}!}} \f{\Omega_1\Omega_{-1}}{\Delta}.
+  \sqrt{\f{\p{n+2}!}{\p{n-3}!}}~ \f{\Omega_+\Omega_-}{\Delta}.
 \end{align}
-Here $\Omega_m$, $\phi_{jm}$, and $\Delta$ are respectively the drive
-strengths, phases, and detuning that appear in the ``bare'' Raman
-drive in \eqref{eq:raman_bare}.
+Here $\Omega_m$, $\phi_{jm}$, and $\Delta$ are respectively the (real)
+driving amplitudes, phases, and detuning that appear in the ``bare''
+Raman drive in \eqref{eq:raman_bare}, and by choice of gauge we set
+$\phi_{j,0}=0$, or equivalently re-define
+$\phi_{jm}-\phi_{j,0}\to\phi_{jm}$.  If the driving lasers are
+phase-matched with $\phi_{js}\equiv s\phi_j$, then the drive further
+simplifies to
+\begin{align}
+  H_{\t{Raman}}^{(\phi)}
+  = \sum_{\substack{L\le2\\0\le M\le L}} \Omega_{LM} \D_{LM}^{(\phi)},
+\end{align}
+with inhomogeneously rotated collective drive operators
+\begin{align}
+  \D_{LM}^{(\phi)}
+  \equiv \sum_j \sp{\cos\p{M\phi_j} D_{LM}^{(j)}
+    + \sin\p{M\phi_j} D_{L,-M}^{(j)}},
+\end{align}
+and uniform coefficients given by \eqref{eq:O_01_0}, \eqref{eq:O_2_0},
+and
+\begin{align}
+  \Omega_{1,-1} = \Omega_{2,-1} = \Omega_{2,-2} \equiv 0,
+  &&
+  \Omega_{L,1} \equiv \sum_{s\in\set{\pm1}} \omega_{Ls},
+  &&
+  \Omega_{2,2} \equiv \omega_{2,2}.
+\end{align}
 

### 2019-10-03 collective_sun/sun_notes.tex
some rearranging text and additional commentary

@@
 this drive spatially homogenous via the gauge transformation
-$c_{j\mu}^\dag \to e^{i\mu\phi_j} c_{j\mu}^\dag$, after which the
-drive becomes simply
+$c_{j\mu}^\dag \to e^{i\mu\phi_j} c_{j\mu}^\dag$, under which
 \begin{align}
-  H_{\t{drive}} = \Omega \D_{1,1}.
+  H_{\t{drive}}^{(\phi)} \to H_{\t{drive}} = \Omega \D_{1,1}.
 \end{align}
@@
 \begin{align}
-  H_{\t{lat}}^{(\phi)} = -J \sum_{q,L} B_L\p{q,\phi} D_{L,0}^{(q)},
+  H_{\t{lat}}^{(\phi)} = -J \sum_{q,L} B_L^{(\phi,q)} D_{L,0}^{(q)},
   \label{eq:H_lat_drive}
 \end{align}
-with effective driving field strengths
+with effective inhomogeneous driving field strengths
 \begin{align}
-  B_L\p{q,\phi} = f_L\p{qa} \sqrt{\f{L+1/2}{n}}
+  B_L^{(\phi,q)} = f_L\p{qa} \sqrt{\f{L+1/2}{n}}
   \sum_\mu \bk{I\mu,L,0;I\mu} f_L\p{\mu\phi},
@@
   H_{\t{Raman},0}^{(\phi)}
-  &\equiv -\sum_j\sum_m \f{\Omega_m^2}{\Delta}
+  &= -\sum_j\sum_m \f{\Omega_m^2}{\Delta}
   {T_{1,m}^{(j)}}^\dag T_{1,m}^{(j)}, & \label{eq:raman_0} \\
   H_{\t{Raman},1}^{(\phi)}
-  &\equiv -\sum_j\sum_{s\in\set{\pm1}}
+  &= -\sum_j\sum_{s\in\set{\pm1}}
   \f{\Omega_0\Omega_s}{\Delta} e^{-i\phi_{js}}
@@
   H_{\t{Raman},2}^{(\phi)}
-  &\equiv -\f{\Omega_+\Omega_-}{\Delta}
+  &= -\f{\Omega_+\Omega_-}{\Delta}
   \sum_j\sum_{s\in\set{\pm1}} e^{-is\p{\phi_{j+}-\phi_{j-}}}
@@
 Appendix \ref{sec:drive_raman_coeff}.  If the driving lasers are
-phase-matched with $\phi_{js}\equiv s\phi_j$, then the drive further
-simplifies to
+phase-matched with $\phi_{js}=s\phi_j$, such that a particle on site
+$j$ acquires a phase $\mp\phi_j$ for every nuclear spin transition
+$\mu\to\mu\pm1$, then the drive $H_{\t{Raman}}^{(\phi)}$ takes the
+form
 \begin{align}
-  H_{\t{Raman}}^{(\phi)}
-  = \sum_{\substack{L\le2\\0\le M\le L}} \Omega_{LM} \D_{LM}^{(\phi)}
+  \left. H_{\t{Raman}}^{(\phi)} \right|_{\phi_{js}=s\phi_j}
+  = \sum_{\substack{L\le2\\0\le M\le L}} \Omega_{LM} \D_{LM}^{(\phi)},
   \label{eq:drive_raman_matched}
@@
 with rotated collective drive operators $\D_{LM}^{(\phi)}$ defined in
-\eqref{eq:drive_rot}, and uniform coefficients $\Omega_{LM}$ that have
-the scaling
+\eqref{eq:drive_rot} and uniform driving amplitudes $\Omega_{LM}$
+provided in Appendix \ref{sec:drive_raman_coeff}.  Similarly to the
+case of a direct drive considered in Section \ref{sec:drive_SOC}, the
+Raman drive $H_{\t{Raman}}^{(\phi)}$ in \eqref{eq:drive_raman_matched}
+can be made homogenous through the gauge transformation
+$c_{j\mu}^\dag \to e^{i\mu\phi_j} c_{j\mu}^\dag$, under which
+\begin{align}
+  \left. H_{\t{Raman}}^{(\phi)} \right|_{\phi_{js}=s\phi_j}
+  \to H_{\t{Raman}}
+  = \sum_{\substack{L\le2\\0\le M\le L}} \Omega_{LM} \D_{LM}.
+\end{align}
+The effective driving amplitudes $\Omega_{LM}$ can be tuned through
+the bare driving amplitudes $\Omega_m$, and satisfy
 \begin{align}
@@
 \end{align}
-Similarly to the case of a direct drive considered in Section
-\ref{sec:drive_SOC}, the Raman drive $H_{\t{Raman}}^{(\phi)}$ in
-\eqref{eq:drive_raman_matched} can be made homogenous through the
-gauge transformation $c_{j\mu}^\dag \to e^{i\mu\phi_j} c_{j\mu}^\dag$,
-under which
-\begin{align}
-  H_{\t{Raman}}^{(\phi)} \to H_{\t{Raman}}
-  \equiv \sum_{\substack{L\le2\\0\le M\le L}} \Omega_{LM} \D_{LM}.
-  \label{eq:drive_raman_homo}
-\end{align}
+In particular, the amplitudes $\Omega_{1,0}$, $\Omega_{2,0}$, and
+$\Omega_{2,1}$ can be simultaneously tuned to zero by setting
+$\Omega_0=\Omega_+=\Omega_-$, at which point the only remaining
+effective drives in $H_{\t{Raman}}$ are $\sim\D_{1,1}$ and
+$\sim\D_{2,2}$ (as $\D_{0,0}\propto\1$ is just constant shift in
+energy).
 
@@
   \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
-  = \sum_L B_L\p{q,\phi} D_{L,0},
+  = \sum_L B_L^{(\phi,q)} D_{L,0},
   \label{eq:compact_SOC_drive}
 \end{align}
-with effective driving field strengths
+with effective inhomogeneous driving field strengths
 \begin{align}
-  B_L\p{q,\phi}
+  B_L^{(\phi,q)}
   \equiv f_L\p{qa} \f12
@@
 \begin{align}
-  H_{\t{lat}}^{(\phi)} = -J \sum_{q,L} B_L\p{q,\phi} D_{L,0}^{(q)}.
+  H_{\t{lat}}^{(\phi)} = -J \sum_{q,L} B_L^{(\phi,q)} D_{L,0}^{(q)}.
 \end{align}
@@
 $\phi_{jm}-\phi_{j,0}\to\phi_{jm}$.  If the driving lasers are
-phase-matched with $\phi_{js}\equiv s\phi_j$, then the drive further
-simplifies to
+phase-matched with $\phi_{js}=s\phi_j$, such that a particle on site
+$j$ acquires a phase $\mp\phi_j$ for every nuclear spin transition
+$\mu\to\mu\pm1$, then the drive $H_{\t{Raman}}^{(\phi)}$ takes the
+form
 \begin{align}
-  H_{\t{Raman}}^{(\phi)}
+  \left. H_{\t{Raman}}^{(\phi)} \right|_{\phi_{js}=s\phi_j}
   = \sum_{\substack{L\le2\\0\le M\le L}} \Omega_{LM} \D_{LM}^{(\phi)},

### 2019-10-03 collective_sun/sun_notes.tex
added a comment on an amplitude-matching drive condition
added comment about simulating large-spin synthetic gauge fields

@@
 this drive spatially homogenous via the gauge transformation
-$c_{j\mu}^\dag \to e^{i\mu\phi_j} c_{j\mu}^\dag$, under which
+$c_{j\mu}^\dag \to e^{i\mu\phi_j} c_{j\mu}^\dag$, which takes
 \begin{align}
@@
 \end{align}
+Spin-orbit coupling with multi-level systems thus allows for the
+simulation of synthetic gauge fields with large spin ($L>1$).  As we
+will see in the following section, multi-laser drives can also
+simulate large-spin synthetic fields with non-zero spin projection
+($M\ne0$) onto a quantization axis.
 
@@
 can be made homogenous through the gauge transformation
-$c_{j\mu}^\dag \to e^{i\mu\phi_j} c_{j\mu}^\dag$, under which
+$c_{j\mu}^\dag \to e^{i\mu\phi_j} c_{j\mu}^\dag$, which takes
 \begin{align}
@@
 $\sim\D_{2,2}$ (as $\D_{0,0}\propto\1$ is just constant shift in
-energy).
+energy).  If instead $\Omega_0=\pm\Omega_\pm$, then the amplitudes
+$\Omega_{1,0}$, $\Omega_{2,0}$ and $\Omega_{1,1}$ vanish, leaving only
+the drives $\D_{2,1}$ and $\D_{2,2}$.
 

### 2019-10-03 collective_sun/sun_notes.tex
some cleanup in notations and conventions

@@
 \end{align}
-In the following sections, we identify various bases for the space of
-operators on the Hilbert space of an $n$-level system.  For brevity,
-we denote this space of operators by $\B_n$, and note that $\B_n$ is
-itself an $n^2$-dimensional Hilbert space equipped with a trace
-(Hilbert-Schmidt) inner product
+In the remainder of this section, we identify various bases for the
+space of operators on the Hilbert space of an $n$-level system.  For
+brevity, we denote this space of operators by $\B_n$, and note that
+$\B_n$ is itself an $n^2$-dimensional Hilbert space equipped with a
+trace (Hilbert-Schmidt) inner product
 \begin{align}
@@
   = \Omega \sum_j
-  \p{e^{-i\phi_j} c_{j,\up}^\dag c_{j,\dn}
-    + e^{i\phi_j} c_{j,\dn}^\dag c_{j,\up}},
+  \p{e^{-i\phi j} c_{j,\up}^\dag c_{j,\dn}
+    + e^{i\phi j} c_{j,\dn}^\dag c_{j,\up}},
   \label{eq:H_drive_2}
 \end{align}
-where $\Omega$ is a (real) driving amplitude, $j$ indexes a lattice
-site, and $\phi_j$ is a site-dependent phase.  The multi-level
-analogue of \eqref{eq:H_drive_2} is
+where $\Omega$ is a (real) driving amplitude, $j\in\mathbb{Z}$ indexes
+a lattice site, and $\phi$ is the relative phase of the drive between
+adjacent lattice sites.  The multi-level analogue of
+\eqref{eq:H_drive_2} is
 \begin{align}
   H_{\t{drive}}^{(\phi)}
-  = -\Omega \sum_j \p{e^{-i\phi_j} T_{1,1}^{(j)}
-    + e^{i\phi_j} {T_{1,1}^{(j)}}^\dag}
+  = -\Omega \sum_j \p{e^{-i\phi j} T_{1,1}^{(j)}
+    + e^{i\phi j} {T_{1,1}^{(j)}}^\dag}
   = \Omega \D_{1,1}^{(\phi)},
@@
   \D_{LM}^{(\phi)}
-  \equiv \sum_j \sp{\cos\p{M\phi_j} D_{LM}^{(j)}
-    + \sin\p{M\phi_j} D_{L,-M}^{(j)}},
+  \equiv \sum_j \sp{\cos\p{M\phi j} D_{LM}^{(j)}
+    + \sin\p{M\phi j} D_{L,-M}^{(j)}},
   \label{eq:drive_rot}
@@
 this drive spatially homogenous via the gauge transformation
-$c_{j\mu}^\dag \to e^{i\mu\phi_j} c_{j\mu}^\dag$, which takes
+$c_{j\mu}^\dag \to e^{i\mu\phi j} c_{j\mu}^\dag$, which takes
 \begin{align}
@@
 invariant under this gauge transformation, but the single-particle
-Hamiltonian in \eqref{eq:H_lat} transforms as
+Hamiltonian in \eqref{eq:H_lat} transforms as (see Appendix
+\ref{sec:lat_drive})
 \begin{align}
@@
   \to H_{\t{lat}}^{(\phi)}
-  = -J \sum_{q,\mu} \cos\p{qa-\mu\phi} c_{q\mu}^\dag c_{q\mu},
+  = -J \sum_{q,\mu} \cos\p{qa-\mu\phi} c_{q\mu}^\dag c_{q\mu}
+  = \sum_{q,L} B_L^{(\phi,q)} D_{L,0}^{(q)},
   \label{eq:H_lat_SOC}
 \end{align}
-where $\phi$ is the relative phase of the driving laser between
-adjacent lattice sites.  In the basis of drive operators, this
-Hamiltonian takes the form (see Appendix \ref{sec:lat_drive})
+with effective inhomogeneous driving field amplitudes
 \begin{align}
-  H_{\t{lat}}^{(\phi)} = -J \sum_{q,L} B_L^{(\phi,q)} D_{L,0}^{(q)},
-  \label{eq:H_lat_drive}
-\end{align}
-with effective inhomogeneous driving field strengths
-\begin{align}
-  B_L^{(\phi,q)} = f_L\p{qa} \sqrt{\f{L+1/2}{n}}
+  B_L^{(\phi,q)} = -J f_L\p{qa} \sqrt{\f{L+1/2}{n}}
   \sum_\mu \bk{I\mu,L,0;I\mu} f_L\p{\mu\phi},
@@
 will see in the following section, multi-laser drives can also
-simulate large-spin synthetic fields with non-zero spin projection
+simulate synthetic large-spin fields with non-zero spin projection
 ($M\ne0$) onto a quantization axis.
@@
   H_{\t{Raman,bare}}^{(\phi)}
-  = \sum_j \sum_{m\in\set{1,0,-1}} \p{\Omega_m e^{-i\phi_{jm}}
+  = \sum_j \sum_{m\in\set{1,0,-1}} \p{\Omega_m e^{-i\phi_m j}
     T_{1,m}^{(j)} \otimes \op{\up}{\dn} + \t{h.c.}}
@@
 where $j$ indexes a lattice site, $\Omega_m$ is a (real) driving
-amplitude; $\phi_{jm}$ is a site- and drive-dependent phase;
-$\op{r}{s}$ for $r,s\in\set{\up,\dn}$ is a pseudo-spin operator for
-the auxiliary degree of freedom off-resonantly addressed by the
-lasers; $\Delta$ is a detuning of the lasers from the energy of an
-excited state; and $\1$ is the identity operator.  If the detuning is
-large, $\Delta\gg\Omega_m$, then at second order in perturbation
-theory we can find that the effective action of
-$H_{\t{Raman,bare}}^{(\phi)}$ within the $\ket{\dn}$-state manifold is
+amplitude; $\phi_m$ is the relative phase of drive $m$ between
+adjacent lattice sites; $\op{r}{s}$ for $r,s\in\set{\up,\dn}$ is a
+pseudo-spin operator for the auxiliary degree of freedom
+off-resonantly addressed by the lasers; $\Delta$ is a detuning of the
+lasers from the energy of an excited state; and $\1$ is the identity
+operator.  If the detuning is large, $\Delta\gg\Omega_m$, then at
+second order in perturbation theory we can find that the effective
+action of $H_{\t{Raman,bare}}^{(\phi)}$ within the $\ket{\dn}$-state
+manifold is
 \begin{align}
@@
   &= -\sum_j\sum_{s\in\set{\pm1}}
-  \f{\Omega_0\Omega_s}{\Delta} e^{-i\phi_{js}}
+  \f{\Omega_0\Omega_s}{\Delta} e^{-i\phi_s j}
   {T_{1,0}^{(j)}}^\dag T_{1,s}^{(j)} + \t{h.c.},
@@
   &= -\f{\Omega_+\Omega_-}{\Delta}
-  \sum_j\sum_{s\in\set{\pm1}} e^{-is\p{\phi_{j+}-\phi_{j-}}}
+  \sum_j\sum_{s\in\set{\pm1}} e^{-is\p{\phi_+-\phi_-}j}
   {T_{1,-s}^{(j)}}^\dag T_{1,s}^{(j)},
@@
 \end{align}
-where by choice of gauge we set $\phi_{j,0}=0$, or equivalently
-re-define $\phi_{jm}-\phi_{j,0}\to\phi_{jm}$, and we identify the
-subscripts $\pm\leftrightarrow\pm1$.  In terms of the drive operators,
-we can expand
+where by choice of gauge we set $\phi_0=0$, or equivalently re-define
+$\phi_m-\phi_0\to\phi_m$, and we identify the subscripts
+$\set{\pm}\leftrightarrow\set{\pm1}$.  In terms of the drive
+operators, we can expand
 \begin{align}
@@
 Appendix \ref{sec:drive_raman_coeff}.  If the driving lasers are
-phase-matched with $\phi_{js}=s\phi_j$, such that a particle on site
-$j$ acquires a phase $\mp\phi_j$ for every nuclear spin transition
+phase-matched with $\phi_s=s\phi$, such that a particle on site $j$
+acquires a phase $\mp\phi j$ for every nuclear spin transition
 $\mu\to\mu\pm1$, then the drive $H_{\t{Raman}}^{(\phi)}$ takes the
@@
 \begin{align}
-  \left. H_{\t{Raman}}^{(\phi)} \right|_{\phi_{js}=s\phi_j}
+  \left. H_{\t{Raman}}^{(\phi)} \right|_{\phi_s=s\phi}
   = \sum_{\substack{L\le2\\0\le M\le L}} \Omega_{LM} \D_{LM}^{(\phi)},
@@
 can be made homogenous through the gauge transformation
-$c_{j\mu}^\dag \to e^{i\mu\phi_j} c_{j\mu}^\dag$, which takes
+$c_{j\mu}^\dag \to e^{i\mu\phi j} c_{j\mu}^\dag$, which takes
 \begin{align}
-  \left. H_{\t{Raman}}^{(\phi)} \right|_{\phi_{js}=s\phi_j}
+  \left. H_{\t{Raman}}^{(\phi)} \right|_{\phi_s=s\phi}
   \to H_{\t{Raman}}
@@
 
-Here we take the single-particle spin-orbit coupled Hamiltonian
-$H_{\t{lat}}^{(\phi)}$ in \eqref{eq:H_lat_SOC}, and derive its
-expression in \eqref{eq:H_lat_drive} in terms of the drive operators
-$D_{LM}$.  In the basis of ordinary spin operators $s_{\mu\mu}$, this
-Hamiltonian takes the form
+Here we derive the drive-operator ($D_{LM}$) expansion of the
+spin-orbit coupled Hamiltonian $H_{\t{lat}}^{(\phi)}$ in
+\eqref{eq:H_lat_SOC}.  In the basis of spin operators
+$s_{\mu\mu}^{(q)} \equiv c_{q\mu}^\dag c_{q\mu}$, this Hamiltonian
+takes the form
 \begin{align}
@@
 \begin{align}
-  w_{\phi,+} \equiv \sum_\mu \cos\p{\mu\phi} s_{\mu\mu},
+  w_{\phi+} \equiv \sum_\mu \cos\p{\mu\phi} s_{\mu\mu},
   &&
-  w_{\phi,-} \equiv \sum_\mu \sin\p{\mu\phi} s_{\mu\mu},
+  w_{\phi-} \equiv \sum_\mu \sin\p{\mu\phi} s_{\mu\mu},
 \end{align}
 which are respectively even and odd under spin inversion
-($\mu\to-\mu$), we can simplify
+($\mu\to-\mu$), we can expand
 \begin{align}
   \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
-  = \cos\p{qa} w_{\phi,+} + \sin\p{qa} w_{\phi,-},
+  = \cos\p{qa} w_{\phi+} + \sin\p{qa} w_{\phi-},
 \end{align}
@@
   \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
-  = \cos\p{qa} \f12 \sum_{L,M} \obk{w_{\phi,+}|D_{LM}} D_{LM}
-  + \sin\p{qa} \f12 \sum_{L,M} \obk{w_{\phi,-}|D_{LM}} D_{LM}.
+  = \cos\p{qa} \f12 \sum_{L,M} \obk{w_{\phi+}|D_{LM}} D_{LM}
+  + \sin\p{qa} \f12 \sum_{L,M} \obk{w_{\phi-}|D_{LM}} D_{LM}.
 \end{align}
@@
 \begin{align}
-  \sum_{L,M} \obk{w_{\phi,+}|D_{LM}} D_{LM}
-  = \sum_{L\,\t{even}} \obk{w_{\phi,+}|D_{L,0}} D_{L,0},
+  \sum_{L,M} \obk{w_{\phi+}|D_{LM}} D_{LM}
+  = \sum_{L\,\t{even}} \obk{w_{\phi+}|D_{L,0}} D_{L,0},
 \end{align}
-and likewise with $w_{\phi,-}$ and odd $L$.  In total, we can
+and likewise with $w_{\phi-}$ and odd $L$.  In total, we can
 write
@@
   = \cos\p{qa} \f12 \sum_{L\,\t{even}}
-  \obk{w_{\phi,+}|D_{L,0}} D_{L,0}
+  \obk{w_{\phi+}|D_{L,0}} D_{L,0}
   + \sin\p{qa} \f12 \sum_{L\,\t{odd}}
@@
   \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
-  = \sum_L B_L^{(\phi,q)} D_{L,0},
+  = \sum_L A_L^{(\phi,q)} D_{L,0},
   \label{eq:compact_SOC_drive}
 \end{align}
-with effective inhomogeneous driving field strengths
+with effective inhomogeneous driving field amplitudes
 \begin{align}
-  B_L^{(\phi,q)}
+  A_L^{(\phi,q)}
   \equiv f_L\p{qa} \f12
@@
 \begin{align}
-  H_{\t{lat}}^{(\phi)} = -J \sum_{q,L} B_L^{(\phi,q)} D_{L,0}^{(q)}.
+  H_{\t{lat}}^{(\phi)} = \sum_{q,L} B_L^{(\phi,q)} D_{L,0}^{(q)},
+  &&
+  B_L^{(\phi,q)} \equiv -J A_L^{(\phi,q)}.
 \end{align}
@@
   \Omega_{L,1}^{(\phi,j)}
-  \equiv \sum_{s\in\set{\pm1}} \omega_{Ls} \cos\phi_{js},
+  \equiv \sum_{s\in\set{\pm1}} \omega_{Ls} \cos\p{\phi_s j},
   &&
   \Omega_{L,-1}^{(\phi,j)}
-  \equiv \sum_{s\in\set{\pm1}} s \omega_{Ls} \sin\phi_{js},
+  \equiv \sum_{s\in\set{\pm1}} s \omega_{Ls} \sin\p{\phi_s j},
 \end{align}
@@
   \Omega_{2,2}^{(\phi,j)}
-  \equiv \omega_{2,2} \cos\p{\phi_{j+}-\phi_{j-}},
+  \equiv \omega_{2,2} \cos\p{\sp{\phi_+-\phi_-}j},
   &&
   \Omega_{2,-2}^{(\phi,j)}
-  \equiv \omega_{2,2} \sin\p{\phi_{j+}-\phi_{j-}},
+  \equiv \omega_{2,2} \sin\p{\sp{\phi_+-\phi_-}j},
 \end{align}
@@
 \end{align}
-Here $\Omega_m$, $\phi_{jm}$, and $\Delta$ are respectively the (real)
+Here $\Omega_m$, $\phi_m$, and $\Delta$ are respectively the (real)
 driving amplitudes, phases, and detuning that appear in the ``bare''
 Raman drive in \eqref{eq:raman_bare}, and by choice of gauge we set
-$\phi_{j,0}=0$, or equivalently re-define
-$\phi_{jm}-\phi_{j,0}\to\phi_{jm}$.  If the driving lasers are
-phase-matched with $\phi_{js}=s\phi_j$, such that a particle on site
-$j$ acquires a phase $\mp\phi_j$ for every nuclear spin transition
-$\mu\to\mu\pm1$, then the drive $H_{\t{Raman}}^{(\phi)}$ takes the
-form
+$\phi_-=0$, or equivalently re-define $\phi_m-\phi_0\to\phi_m$.  If
+the driving lasers are phase-matched with $\phi_s=s\phi$, such that a
+particle on site $j$ acquires a phase $\mp\phi j$ for every nuclear
+spin transition $\mu\to\mu\pm1$, then the drive
+$H_{\t{Raman}}^{(\phi)}$ takes the form
 \begin{align}
-  \left. H_{\t{Raman}}^{(\phi)} \right|_{\phi_{js}=s\phi_j}
+  \left. H_{\t{Raman}}^{(\phi)} \right|_{\phi_s=s\phi}
   = \sum_{\substack{L\le2\\0\le M\le L}} \Omega_{LM} \D_{LM}^{(\phi)},
@@
   \D_{LM}^{(\phi)}
-  \equiv \sum_j \sp{\cos\p{M\phi_j} D_{LM}^{(j)}
-    + \sin\p{M\phi_j} D_{L,-M}^{(j)}},
+  \equiv \sum_j \sp{\cos\p{M\phi j} D_{LM}^{(j)}
+    + \sin\p{M\phi j} D_{L,-M}^{(j)}},
 \end{align}

### 2019-10-04 collective_sun/sun_notes.tex
changed some notation and added first results for weak SOC

@@
 \begin{align}
-  B_L^{(\phi,q)} = -J f_L\p{qa} \sqrt{\f{L+1/2}{n}}
-  \sum_\mu \bk{I\mu,L,0;I\mu} f_L\p{\mu\phi},
+  B_L^{(\phi,q)} = -J f_L\p{qa} W_L^{(\phi)},
+\end{align}
+where
+\begin{align}
+  W_L^{(\phi)} \equiv \sqrt{\f{L+1/2}{n}}
+  \sum_\mu \bk{I\mu,L,0;I\mu} f_L\p{\mu\phi}
   &&
@@
 ($M\ne0$) onto a quantization axis.
+
+In the regime of weak SOC with $\abs{\phi}\ll1$, we find that
+\begin{align}
+  W_L^{(\phi)} = c_L \sqrt{n_L\p{L+1/2}} \phi^L + O\p{\phi^{L+2}},
+  &&
+  n_L \equiv \prod_{\ell=-L}^L\p{n-\ell},
+\end{align}
+for
+\begin{align}
+  c_0 = 1,
+  &&
+  c_1 = \f16,
+  &&
+  c_2 = \f1{30},
+  &&
+  c_3 = \f1{140},
+  &&
+  c_4 = \f1{630},
+  &&
+  c_5 = \f1{2772}.
+\end{align}
 
@@
 and resolve the identity (super-)operator $\I$ with
-$\I w_{\phi,\pm} = w_{\phi,\pm}$ in the basis of the drive
-operators $D_{LM}$, finding (see Appendix \ref{sec:changing_bases})
+$\I w_{\phi\pm} = w_{\phi\pm}$ in the basis of the drive operators
+$D_{LM}$ (see Appendix \ref{sec:changing_bases}), finding
 \begin{align}
@@
   + \sin\p{qa} \f12 \sum_{L\,\t{odd}}
-  \obk{w_{\phi,-}|D_{L,0}} D_{L,0},
+  \obk{w_{\phi-}|D_{L,0}} D_{L,0},
 \end{align}
@@
   \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
-  = \sum_L A_L^{(\phi,q)} D_{L,0},
-  \label{eq:compact_SOC_drive}
-\end{align}
-with effective inhomogeneous driving field amplitudes
-\begin{align}
-  A_L^{(\phi,q)}
-  \equiv f_L\p{qa} \f12
-  \sum_\mu f_L\p{\mu\phi} \obk{s_{\mu\mu}|D_{L,0}}
-  = f_L\p{qa} \sqrt{\f{L+1/2}{n}}
-  \sum_\mu \bk{I\mu,L,0;I\mu} f_L\p{\mu\phi},
-\end{align}
-where
-\begin{align}
+  = \sum_L f_L\p{qa} W_L^{(\phi)} D_{L,0},
+  &&
   f_L\p{\theta} \equiv
@@
     \sin\theta & L~\t{odd}
-  \end{cases}.
+  \end{cases},
+  \label{eq:compact_SOC_drive}
+\end{align}
+with effective driving field amplitudes
+\begin{align}
+  W_L^{(\phi)}
+  \equiv \f12 \obk{w_{\phi,\p{-1}^L}|D_{L,0}}
+  = \f12 \sum_\mu f_L\p{\mu\phi} \obk{s_{\mu\mu}|D_{L,0}}
+  = \sqrt{\f{L+1/2}{n}}
+  \sum_\mu \bk{I\mu,L,0;I\mu} f_L\p{\mu\phi}.
 \end{align}
@@
   &&
-  B_L^{(\phi,q)} \equiv -J A_L^{(\phi,q)}.
+  B_L^{(\phi,q)} \equiv -J f_L\p{qa} W_L^{(\phi)}.
 \end{align}
@@
 \begin{align}
-  \Omega_{0,0} \equiv -\f1{\sqrt{2n}} \sum_m\f{\Omega_m^2}{\Delta},
+  \Omega_{0,0} \equiv -\f1{\sqrt{2}}~ \f1{\sqrt{n}}~
+  \sum_m\f{\Omega_m^2}{\Delta},
   &&
-  \Omega_{1,0} \equiv \sqrt{\f{3}{2n\p{n^2-1}}}~
-  \f{\Omega_+^2 - \Omega_-}{\Delta},
+  \Omega_{1,0} \equiv \sqrt{\f32}~ \f1{\sqrt{n_1}}~
+  \f{\Omega_+^2 - \Omega_-^2}{\Delta},
   \label{eq:O_01_0}
@@
 \begin{align}
-  \Omega_{2,0} \equiv \sqrt{\f{n^2-4}{10n\p{n^2-1}}}~
+  \Omega_{2,0} \equiv \f1{\sqrt{10}}~ \f{\sqrt{n_2}}{n_1}~
   \f{\Omega_+ + \Omega_- - 2\Omega_0}{\Delta},
@@
   \Omega_{L,1}^{(\phi,j)}
-  \equiv \sum_{s\in\set{\pm1}} \omega_{Ls} \cos\p{\phi_s j},
-  &&
+  &\equiv \sum_{s\in\set{\pm1}} \omega_{Ls} \cos\p{\phi_s j},
+  &
   \Omega_{L,-1}^{(\phi,j)}
-  \equiv \sum_{s\in\set{\pm1}} s \omega_{Ls} \sin\p{\phi_s j},
-\end{align}
-\begin{align}
+  &\equiv \sum_{s\in\set{\pm1}} s \omega_{Ls} \sin\p{\phi_s j}, \\
   \Omega_{2,2}^{(\phi,j)}
-  \equiv \omega_{2,2} \cos\p{\sp{\phi_+-\phi_-}j},
-  &&
+  &\equiv \omega_{2,2} \cos\p{\sp{\phi_+-\phi_-}j},
+  &
   \Omega_{2,-2}^{(\phi,j)}
-  \equiv \omega_{2,2} \sin\p{\sp{\phi_+-\phi_-}j},
-\end{align}
-\begin{align}
-  \omega_{1\pm} \equiv \f{\sqrt3}{\sqrt{n\p{n^2-1}}}~
+  &\equiv \omega_{2,2} \sin\p{\sp{\phi_+-\phi_-}j}, \\
+  \omega_{1\pm}
+  &\equiv \sqrt{3} ~\f1{\sqrt{n_1}}~
   \f{\Omega_0 \Omega_\pm}{\Delta},
-  &&
-  \omega_{2\pm} \equiv \pm \f{\sqrt{3/5}}{n\p{n^2-1}}
-  \sqrt{\f{\p{n+2}!}{\p{n-3}!}}~ \f{\Omega_0\Omega_\pm}{\Delta},
-\end{align}
-\begin{align}
-  \omega_{2,2} \equiv -\f{\sqrt{6/5}}{n\p{n^2-1}}
-  \sqrt{\f{\p{n+2}!}{\p{n-3}!}}~ \f{\Omega_+\Omega_-}{\Delta}.
+  &
+  \omega_{2\pm}
+  &\equiv \pm \sqrt{\f35}~ \f{\sqrt{n_2}}{n_1}~
+  \f{\Omega_0\Omega_\pm}{\Delta}, \\
+  \omega_{2,2}
+  &\equiv -\sqrt{\f65}~ \f{\sqrt{n_2}}{n_1}~
+  \f{\Omega_+\Omega_-}{\Delta},
+  &
+  n_L
+  &\equiv \prod_{\ell=-L}^L \p{n-\ell}.
 \end{align}

### 2019-10-04 multi_body_interactions/papers/theory/paper.tex
minor cleanup

@@
 suitable for the phenomenological description of these interactions,
-and particular for our eventual development of a low-energy effective
-theory.
+and in particular for our eventual development of a low-energy
+effective theory.
 

### 2019-10-04 soc_squeezing/collective_paper/collective_spin.tex
add graphics path (which was removed accidentally for submission)
minor cleanup

@@
 \newcommand{\sref}[1]{\protect\subref{#1}} % for referencing subfigures
+\graphicspath{{./figures/}} % set path for all figures
 
@@
 TST expansion is essentially perturbative in the time $t$, which
-implies that the validity of this a formal expansion eventually breaks
+implies that its validity as a formal expansion eventually breaks
 down.  Precisely characterizing the implications of these last two
@@
 $\rho_S^{(2)}\equiv\rho_S^{(1)}+\O_S$ with $\O_S$ any traceless
-operator on $S$ with squared norm
-$\norm{\O_S}^2\equiv\tr\p{\O_S^\dag\O_S}\le\tr\1_S$, then by linearity
-we find that
+operator on $S$ with operator norm $\norm{\O_S}\le1/\tr\1_S$
+(i.e.~such that $\rho_S^{(2)}$ remains positive semi-definite, or a
+valid quantum state), then by linearity we find that
 \begin{align}

### 2019-10-04 soc_squeezing/soc_squeezing_notes.tex
changed figure direcotry and set S \equiv N/2
minor cleanup

@@
 \usepackage[caption=false]{subfig} % subfigures (via \subfloat[]{})
-\graphicspath{{./figures/}} % set path for all figures
+\graphicspath{{./figures/squeezing/}} % set path for all figures
 
@@
 energies.  By solving this single-particle energy conservation
-condition, one finds that we must have $\set{p,q}\ne\set{r,s}$; that
-is, atoms in different clock states can only interact via
+condition, one finds that we must have $\set{p,q}=\set{r,s}$; that is,
+atoms in different clock states can only interact via
 \begin{enumerate*}[label=(\roman*)]
@@
 the corresponding energies are
-$\bk{Sm|H_{\t{int}}|S,m}=-\p{U/L}S\p{S+1}$.  In particular, the
+$\bk{Sm|H_{\t{int}}|Sm}=-\p{U/L}S\p{S+1}$.  In particular, the
 ground-state $S=N/2$ manifold is spanned by the Dicke states
@@
 and we consider $N$-particle states initially in the ground-state
-manifold $\G_0$ of $H_0$, which have total spin $S=N/2$.  If the
+manifold $\G_0$ of $H_0$, which have total spin $S\equiv N/2$.  If the
 largest eigenvalue of $V$ is smaller in magnitude than half of the
@@
 \begin{align}
-  H_{\t{eff}}
+  H_{\t{eff}}^{(2)}
   = -\f12 \P_0 \p{\sp{\O V, \I V \P_0} - \sp{\O V, \P_0 V \I}} \P_0

### 2019-10-04 soc_squeezing/squeezing_paper/supplementary.tex
minor cleanup

@@
 \includegraphics[width=0.5\textwidth]{relative_error.pdf}
-\caption{{\bf Relative error} between maximal squeezing (measured in dB) obtained by the OAT [Eqn.~(3)] and spin [Eqn.~(2)] models of the main text in a system of 20 particles.
+\caption{{\bf Relative error} between maximal squeezing (measured in dB) obtained by the OAT [Eqn.~(4)] and spin [Eqn.~(3)] models of the main text in a system of 20 particles.
 The OAT model correctly captures the maximal squeezing (in dB) of the spin model to within 3\% (marked by the horizontal reference line) within the gap-protected regime $\widetilde{B}/U<0.06$.
@@
 \caption{{\bf Comparison between the OAT and the spin model in the presence of decoherence.}
-({\bf a}) The difference between the maximal squeezing (measured in dB) obtained by the OAT [Eqn.~(3)] and spin [Eqn.~(2)] models increases with the particle number $N$ and the single-particle spontaneous emission rate $\gamma$.
+({\bf a}) The difference between the maximal squeezing (measured in dB) obtained by the OAT [Eqn.~(4)] and spin [Eqn.~(3)] models increases with the particle number $N$ and the single-particle spontaneous emission rate $\gamma$.
 This disagreement is attributed in part to the fact that spontaneous emission transfers population of the collective spin state outside of the Dicke manifold, violating an assumption of the OAT model; see panel ({\bf b}).

### 2019-10-04 soc_squeezing/scripts/correlator_methods.py

@@ -723,15 +723,18 @@ def deriv_op_vec_to_vals(bare_deriv_op_vec, spin_num, init_state,
     for deriv_op, order in deriv_op_vec.keys():
         for op, val in deriv_op_vec[deriv_op,order].items():
 
+            # initial magnitude / phase of this operator, assuming we have already computed it
             init_ln_val = init_ln_vals.get(op)
+
+            # if we can't acquire the initial magnitude / phase, compute it
             if init_ln_val is None:
                 init_ln_val = init_ln_val_func(op)
-                if init_ln_val is None: continue
+                if init_ln_val is None: continue # the initial magnitude is 0
                 init_ln_vals[op] = init_ln_val
 
             term_ln_mag = np.log(abs(val)) + init_ln_val[0]
-            term = np.exp(term_ln_mag) * init_ln_val[1] * np.exp(1j*np.angle(val))
-            deriv_vals[deriv_op][order] += term
+            term_phase = init_ln_val[1] * np.exp(1j*np.angle(val))
+            deriv_vals[deriv_op][order] += term_phase * np.exp(term_ln_mag)
 
     return deriv_vals
 

### 2019-10-04 soc_squeezing/squeezing_paper/main.tex
pushing final version of squeezing paper submitted to PRR

@@
   &\qquad + U \sum_j \hat n_{j,\up} \hat n_{j,\dn},
-\label{eq:FermiHubbard}
 \end{align}
@@
   + \cos(qa) \hat n_{q,\dn}}.
-\label{eq:FermiHubbards}
+\label{eq:FermiHubbard}
 \end{align}
@@
   \times N/\abs*{\braket{\hat{\v S}}}^2,
+  \label{eq:sqz}
 \end{align}
@@
 Further improvements upon this scheme can be made by introducing a time-dependent driving field that transforms the OAT Hamiltonian into a two-axis twisting (TAT) one.
-While the OAT model initially generates squeezing faster than the TAT model, the squeezing generation rate of OAT (measured in dB per second) falls off with time, while the squeezing generation rate for TAT remains approximately constant until reaching Heisenberg-limited amount of spin squeezing with $\xi^2_{\t{opt}}\sim N^{-1}$\cite{kitagawa1993squeezed, supplementary_material}.
+While the OAT model initially generates squeezing faster than the TAT model, the squeezing generation rate of OAT (measured in dB per second) falls off with time, while the squeezing generation rate for TAT remains approximately constant until reaching Heisenberg-limited amount of spin squeezing with $\xi^2_{\t{opt}}\sim N^{-1}$ (see Appendix \ref{sec:time_series})\cite{kitagawa1993squeezed}.
 
@@
 The maximally attainable amount of spin squeezing by each protocol in Fig.~\ref{fig:optima_coherent} depends only on the atom number $N$, while the shortest attainable time is determined by choosing the largest SOC angles $\phi_\x=\phi_\y\equiv\phi$ which saturate $\widetilde{B}/U\approx0.05$.
-We impose this constraint on $\widetilde{B}/U$ to ensure validity of the OAT Hamiltonian perturbatively derived in Appendix \ref{sec:derivation_OAT}\cite{supplementary_material}.
+We impose this constraint on $\widetilde{B}/U$ to ensure validity of the OAT Hamiltonian perturbatively derived in Appendix \ref{sec:derivation_OAT} (see also Appendix \ref{sec:benchmarking}).
 
@@
 The decoherence considered in this work also limits maximally achievable squeezing to $\sim20$ dB less than the decoherence-free maxima shown in Fig.~\ref{fig:optima_coherent}.
-Sample plots of squeezing over time for particular choices of lattice size ($\ell$) and depth ($V_0/E_{\t{R}}$) are provided in the Supplementary Material\cite{supplementary_material}.
+Sample plots of squeezing over time for particular choices of lattice size ($\ell$) and depth ($V_0/E_{\t{R}}$) are provided in Appendix \ref{sec:time_series}.
 }
@@
 \end{align}
-We benchmark the validity of this effective Hamiltonian via exact simulations of the spin [Eqn.~\eqref{eq:SpinModel}] and OAT [Eqn.~\eqref{eq:H_eff}] Hamiltonians in a system of 20 spins, finding that the relative error in maximal squeezing (in dB) of the OAT model is less than 3\% when $\widetilde{B}/U<0.06$\cite{supplementary_material}.
+We benchmark the validity of this effective Hamiltonian via exact simulations of the spin [Eqn.~\eqref{eq:SpinModel}] and OAT [Eqn.~\eqref{eq:H_eff}] Hamiltonians in a system of 20 spins, finding that the relative error in maximal squeezing (in dB) of the OAT model is less than 3\% when $\widetilde{B}/U<0.06$ (see Appendix \ref{sec:benchmarking}).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Numerical benchmarking of the OAT model}
+\label{sec:benchmarking}
+\setcounter{figure}{0}
+
+Here we provide additional information about our benchmarking of the one-axis twisting model against the spin model.
+This benchmarking was performed via exact simulations of a 20-spin system.
+Fig.~\ref{fig:relative_error} shows the relative error in maximal squeezing of the OAT model (measured against the spin model) as a function of the reduced field variance $\widetilde{B}/U$.
+Here squeezing is measured in decibels (dB) by $-10\log_{10}\xi^2$ for the squeezing parameter $\xi^2$ define in Eqn.~\eqref{eq:sqz}.
+The relative error in maximal squeezing (in dB) by the OAT model is less than 3\% when $\widetilde{B}/U<0.06$.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.35\textwidth]{relative_error.pdf}
+\caption{{\bf Relative error} between maximal squeezing (measured in dB) obtained by the OAT [Eqn.~\eqref{eq:H_eff}] and spin [Eqn.~\eqref{eq:SpinModel}] models of the main text in a system of 20 particles.
+The OAT model correctly captures the maximal squeezing (in dB) of the spin model to within 3\% (marked by the horizontal reference line) within the gap-protected regime $\widetilde{B}/U<0.06$.
+}
+\label{fig:relative_error}
+\end{figure}
+
+In principle, spin-changing decoherence compromises the validity of the OAT model, as its perturbative derivation in Appendix \ref{sec:derivation_OAT} relies on spin population remaining primarily within the Dicke manifold.
+This assumption breaks down in the presence of, for example, spontaneous emission, which transfers population outside of the Dicke manifold.
+Nonetheless, we find decent agreement between the OAT and spin models when decoherence is sufficiently weak (see Fig.~\ref{fig:benchmarking_decay}).
+
+\begin{figure*}
+\centering
+\includegraphics[width=0.75\textwidth]{benchmarking_decay.pdf}
+\caption{{\bf Comparison between the OAT and the spin model in the presence of decoherence.}
+({\bf a}) The difference between the maximal squeezing (measured in dB) obtained by the OAT [Eqn.~\eqref{eq:H_eff}] and spin [Eqn.~\eqref{eq:SpinModel}] models increases with the particle number $N$ and the single-particle spontaneous emission rate $\gamma$.
+This disagreement is attributed in part to the fact that spontaneous emission transfers population of the collective spin state outside of the Dicke manifold, violating an assumption of the OAT model; see panel ({\bf b}).
+The rate of population transfer outside of the Dicke manifold increases with both particle number and spontaneous emission rate.
+(Parameters for simulations in this figure: $U=1000$ Hz, $J=200$ Hz, and $\phi=\pi/20$).
+}
+\label{fig:benchmarking_decay}
+\end{figure*}
 
@@
 In the language of the section that follows, we would say that this decoherence is captured by the sets of jump operators $\J_-\equiv\set{s_-^{(j)}}$ and $\J_\z\equiv\set{s_\z^{(j)}}$ with corresponding decoherence rates $\gamma_-=\gamma_\z=\Gamma$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Time-series of squeezing via OAT and TAT}
+\label{sec:time_series}
+\setcounter{figure}{0}
+
+Figure \ref{fig:squeezing_example} shows an example of squeezing over time via OAT and TAT, both with and without decoherence via decay and dephasing of individual spins.
+The OAT model initially generates squeezing faster than the TAT model, but the squeezing generation rate of OAT (measured in dB per second) falls off with time.
+The squeezing generation rate for TAT, meanwhile, remains approximately constant (in the absence of decoherence) until squeezing via TAT surpasses that of OAT.
+In the absence of decoherence, OAT achieves a maximal amount of squeezing that scales as $\xi^2\sim N^{-2/3}$, while TAT achieves Heisenberg-limited squeezing with $\xi^2\sim N^{-1}$.
+Note that our method for computing squeezing via TAT in the presence of decoherence (described in Appendix \ref{sec:collective_simulation}) is not capable of computing squeezing for the full range of times shown in Fig.~\ref{fig:squeezing_example}; the corresponding time-series data in this figure is therefore shown up to the point at which this method breaks down.
+
+\begin{figure*}
+\centering
+\includegraphics{squeezing_example_L40.pdf}
+\includegraphics{squeezing_example_L100.pdf}
+\caption{{\bf Squeezing via OAT and TAT} in a 2D section of the 3D $^{87}$Sr optical lattice clock, shown for ({\bf a}) $\ell=40$ and ({\bf b}) $\ell=100$ sites per axis (with $N=\ell^2$ atoms total), and a lattice depth of $V_0=4~E_{\t{R}}$, where $E_{\t{R}}$ is the atomic lattice recoil energy.
+Atoms are confined along the direction transverse to the 2D layer by a lattice of depth 60 $E_{\t{R}}$.
+Squeezing over time is shown for OAT (blue) and TAT (green), both with (solid lines) and without (dashed lines) decoherence via uncorrelated decay and dephasing of individual spins at rates of $0.1~\t{sec}^{-1}$ (see Appendix \ref{sec:decoherence}).
+}
+\label{fig:squeezing_example}
+\end{figure*}
 
@@
 Thus, in this regime, losses are as relevant as light scattering.
-The role of inelastic interactions could be mitigated by either operating at deeper lattices as shown in Fig.
-\ref{fig:inelastic_squeezing}, or by using nuclear spin states to generate the squeezing instead of the clock states directly.
+The role of inelastic interactions could be mitigated by either operating at deeper lattices as shown in Fig.~\ref{fig:inelastic_squeezing}, or by using nuclear spin states to generate the squeezing instead of the clock states directly.
 

### 2019-10-10 collective_sun/sun_notes.tex
change some spacing and fix error in using J <--> I
add conjectured pattern for weak SOC fields
changed convensions in definition of transition operator computed product of drive operators

@@
 
+\DeclareMathOperator{\sign}{sign}
+\DeclareMathOperator{\signs}{signs}
+
 \def\obra#1{\mathinner{({#1}|}}
@@
 \usepackage[inline]{enumitem} % in-line lists and \setlist{} (below)
+\setlist[enumerate,1]{label={(\roman*)}} % default in-line numbering
 \setlist{nolistsep} % more compact spacing between environments
@@
   T_{LM}
-  \equiv \sqrt{\f{2L+1}{2I+1}} \sum_{\mu,\nu}
-  \bk{I\mu,LM;I\nu} \op{\nu}{\mu},
+  \equiv \p{-1}^M \sqrt{\f{2L+1}{2I+1}} \sum_{\mu,\nu}
+  \bk{I\mu;LM|I\nu} \op{\nu}{\mu},
   \label{eq:trans_ops}
@@
 $n$-level nuclear spin onto a quantization axis; and
-$\bk{I\mu;I\nu,LM}$ is a Clebsch-Gordan coefficient.  The transition
-operator $T_{LM}$ is proportional to the transition induced on a
-nuclear spin by the absorption of a spin-$L$ particle with spin
-projection $M$ onto a quantization axis.  Similarly to the GGM
-operators in \eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag}, the
+$\bk{I\mu;LM|I\nu}$ is a Clebsch-Gordan coefficient.  Note that our
+definition of a transition operators differs from that in
+Refs.~\cite{kryszewski2006alternative, bertlmann2008bloch} by the sign
+$\p{-1}^M$; this choice of sign convention will simplify much of our
+work below.
+
+The transition operator $T_{LM}$ is proportional to the transition
+induced on a nuclear spin by the absorption of a spin-$L$ particle
+with spin projection $M$ onto a quantization axis.  Similarly to the
+GGM operators in \eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag}, the
 transition operators in \eqref{eq:trans_ops} provide a complete basis
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Drive operators}
+\subsection{Drive opermators}
 \label{sec:drive_ops}
@@
 \begin{align}
-  D_{L,0} \equiv \sqrt{2}~ T_{L,0},
+  D_{L,0} \equiv T_{L,0},
   &&
-  D_{LM} \stackrel{M>0}{\equiv} -\p{T_{L\abs{M}} + T_{L\abs{M}}^\dag},
+  D_{LM} \stackrel{M\ne0}{\equiv} \f{\sigma_M}{\sqrt{2}}
+  \sp{T_{L\abs{M}} + \sign\p{M}T_{L\abs{M}}^\dag},
   &&
-  D_{LM} \stackrel{M<0}{\equiv} i\p{T_{L\abs{M}} - T_{L\abs{M}}^\dag},
+  \sigma_M \equiv \sign\p{M}^{3/2},
   \label{eq:drive_ops}
@@
 \eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag}, the drive operators
-are equal to the Pauli operators in the case of $n=2$, and satisfy the
-orthonormality condition
+are proportional to the Pauli operators in the case of $n=2$, and
+satisfy the orthonormality condition
 \begin{align}
-  \obk{D_{LM}|D_{L'M'}} = 2\delta_{LL'}\delta_{MM'},
+  \obk{D_{LM}|D_{L'M'}} = \delta_{LL'}\delta_{MM'},
 \end{align}
@@
 \begin{align}
-  H_{\t{int}} = -\f{u}{4} \sum_{p,q} \v D^{(p)}\c\v D^{(q)}
-  = -\f{u}{4} \v\D \c \v\D,
+  H_{\t{int}} = -\f{u}{2} \sum_{p,q} \v D^{(p)}\c\v D^{(q)}
+  = -\f{u}{2} \v\D \c \v\D,
   &&
@@
   \to \p{S_\z,S_\x,S_\y}
-  \times \sqrt{\f{3}{I \p{I+1/2} \p{I+1}}}.
+  \times \f{2\sqrt{3}}{\sqrt{\p{n-1}n\p{n+1}}}.
 \end{align}
@@
 \begin{align}
-  B_L^{(\phi,q)} = -J f_L\p{qa} W_L^{(\phi)},
+  B_L^{(\phi,q)} = -J w_L\p{qa} A_L^{(\phi)},
 \end{align}
@@
 \begin{align}
-  W_L^{(\phi)} \equiv \sqrt{\f{L+1/2}{n}}
-  \sum_\mu \bk{I\mu,L,0;I\mu} f_L\p{\mu\phi}
-  &&
-  f_L\p{\theta} \equiv
+  w_L\p{\theta} \equiv
   \begin{cases}
@@
     \sin\theta & L~\t{odd}
-  \end{cases}.
+  \end{cases},
+  &&
+  A_L^{(\phi)} \equiv \sqrt{\f{2L+1}{n}}
+  \sum_\mu \bk{I\mu;L,0|I\mu} w_L\p{\mu\phi}.
 \end{align}
@@
 
-In the regime of weak SOC with $\abs{\phi}\ll1$, we find that
+In the regime of weak SOC with $\abs{\phi}\ll1$, by considering
+\begin{enumerate*}
+\item all $L\le12$ for arbitrary $n$, and
+\item all $L<n$ for all $n\le25$
+\end{enumerate*}
+(and conjecturing the same pattern for all $L$ and $n$), we find that
+% expand Clebsch-Gordan coefficients in a power series according to
+% \url{functions.wolfram.com/07.38.06.0008.01}
 \begin{align}
-  W_L^{(\phi)} = c_L \sqrt{n_L\p{L+1/2}} \phi^L + O\p{\phi^{L+2}},
+  A_L^{(\phi)} = c_L \sqrt{n_L}\, \phi^L + O\p{\phi^{L+2}},
+  \label{eq:A_L_phi_small}
+\end{align}
+where
+\begin{align}
+  c_L \equiv \p{-1}^{\lfloor L/2\rfloor}
+  \sqrt{2L+1}\, \f{L!}{\p{2L+1}!},
   &&
-  n_L \equiv \prod_{\ell=-L}^L\p{n-\ell},
-\end{align}
-for
-\begin{align}
-  c_0 = 1,
-  &&
-  c_1 = \f16,
-  &&
-  c_2 = \f1{30},
-  &&
-  c_3 = \f1{140},
-  &&
-  c_4 = \f1{630},
-  &&
-  c_5 = \f1{2772}.
+  n_L \equiv \prod_{\ell=-L}^L\p{n-\ell}.
 \end{align}
@@
 adjacent lattice sites; $\op{r}{s}$ for $r,s\in\set{\up,\dn}$ is a
-pseudo-spin operator for the auxiliary degree of freedom
+pseudo-spin operator for the auxiliary degree of freedom that is
 off-resonantly addressed by the lasers; $\Delta$ is a detuning of the
@@
 operator.  If the detuning is large, $\Delta\gg\Omega_m$, then at
-second order in perturbation theory we can find that the effective
-action of $H_{\t{Raman,bare}}^{(\phi)}$ within the $\ket{\dn}$-state
-manifold is
+second order in perturbation theory the effective action of
+$H_{\t{Raman,bare}}^{(\phi)}$ within the $\ket{\dn}$-state manifold
+takes the form
 \begin{align}
@@
 \end{align}
-with
+where $H_{\t{Raman},\tau}^{(\phi)}$ generates a nuclear spin
+transition $\mu\to\mu\pm\tau$:
 \begin{align}
@@
   &= -\sum_j\sum_{s\in\set{\pm1}}
-  \f{\Omega_0\Omega_s}{\Delta} e^{-i\phi_s j}
+  \f{\Omega_0\Omega_s}{\Delta} e^{-i\p{\phi_s-\phi_0} j}
   {T_{1,0}^{(j)}}^\dag T_{1,s}^{(j)} + \t{h.c.},
@@
   H_{\t{Raman},2}^{(\phi)}
-  &= -\f{\Omega_+\Omega_-}{\Delta}
-  \sum_j\sum_{s\in\set{\pm1}} e^{-is\p{\phi_+-\phi_-}j}
-  {T_{1,-s}^{(j)}}^\dag T_{1,s}^{(j)},
+  &= -\sum_j\sum_{s\in\set{\pm1}}
+  \f{\Omega_+\Omega_-}{\Delta} e^{-i\p{\phi_s-\phi_{-s}}j}
+  {T_{1,-s}^{(j)}}^\dag T_{1,s}^{(j)}.
   \label{eq:raman_2}
 \end{align}
-where by choice of gauge we set $\phi_0=0$, or equivalently re-define
-$\phi_m-\phi_0\to\phi_m$, and we identify the subscripts
-$\set{\pm}\leftrightarrow\set{\pm1}$.  In terms of the drive
-operators, we can expand
+In terms of the drive operators, we can expand
 \begin{align}
@@
 Appendix \ref{sec:drive_raman_coeff}.  If the driving lasers are
-phase-matched with $\phi_s=s\phi$, such that a particle on site $j$
-acquires a phase $\mp\phi j$ for every nuclear spin transition
-$\mu\to\mu\pm1$, then the drive $H_{\t{Raman}}^{(\phi)}$ takes the
-form
+phase-matched with $\phi_m=m\phi$, such that a nuclear spin transition
+$\mu\to\mu+m$ on site $j$ imprints the phase $-m\phi j$, then the
+drive $H_{\t{Raman}}^{(\phi)}$ takes the form
 \begin{align}
-  \left. H_{\t{Raman}}^{(\phi)} \right|_{\phi_s=s\phi}
+  \left. H_{\t{Raman}}^{(\phi)} \right|_{\phi_m=m\phi}
   = \sum_{\substack{L\le2\\0\le M\le L}} \Omega_{LM} \D_{LM}^{(\phi)},
@@
 \begin{align}
-  \left. H_{\t{Raman}}^{(\phi)} \right|_{\phi_s=s\phi}
+  \left. H_{\t{Raman}}^{(\phi)} \right|_{\phi_m=m\phi}
   \to H_{\t{Raman}}
@@
 \begin{itemize}
+\item Products and commutators of drive operators
+\item Visualization with generalized Bloch spheres
+  \begin{itemize}
+  \item Single-particle eigenstates of drive operators
+  \item Single-particle dynamics induced by drive operators
+  \end{itemize}
+\item Classifying collective states (e.g.~eigenstates of interactions,
+  spin-wave-like states)
 \item Implications of SOC
   \begin{itemize}
-  \item Perturbative regime: $\phi\ll1$
+  \item Effective collective dynamics for $\abs{\phi}\ll1$
+  \item Deriving \eqref{eq:A_L_phi_small}
   \item Special angles $\phi$
   \end{itemize}
-\item Squeezing
-  \begin{itemize}
-  \item Collection of $n$-level spins
-  \item Generalizations of squeezing beyond SU(2)
-  \end{itemize}
-\item Consider a different type of drive
+\item Generalizations of squeezing beyond SU(2)?
 \item Consider different interactions
@@
   \end{itemize}
+\item Consider different drives
 \end{itemize}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Changing operator bases for collective SU($n$)-symmetric
-  interactions}
+\section{Changing operator bases for SU($n$)-symmetric interactions}
 \label{sec:changing_bases}
@@
 which essentially consists of two-body interactions ($p\ne q$) and
-self-interactions ($p=q$), respectively of the form
+single-body terms ($p=q$), respectively of the form
 \begin{align}
@@
 \begin{align}
-  w_{\phi+} \equiv \sum_\mu \cos\p{\mu\phi} s_{\mu\mu},
+  \tilde d_{\phi+} \equiv \sum_\mu \cos\p{\mu\phi} s_{\mu\mu},
   &&
-  w_{\phi-} \equiv \sum_\mu \sin\p{\mu\phi} s_{\mu\mu},
+  \tilde d_{\phi-} \equiv \sum_\mu \sin\p{\mu\phi} s_{\mu\mu},
 \end{align}
@@
   \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
-  = \cos\p{qa} w_{\phi+} + \sin\p{qa} w_{\phi-},
+  = \cos\p{qa} \tilde d_{\phi+} + \sin\p{qa} \tilde d_{\phi-},
 \end{align}
 and resolve the identity (super-)operator $\I$ with
-$\I w_{\phi\pm} = w_{\phi\pm}$ in the basis of the drive operators
+$\I \tilde d_{\phi\pm} = \tilde d_{\phi\pm}$ in the basis of the drive operators
 $D_{LM}$ (see Appendix \ref{sec:changing_bases}), finding
@@
   \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
-  = \cos\p{qa} \f12 \sum_{L,M} \obk{w_{\phi+}|D_{LM}} D_{LM}
-  + \sin\p{qa} \f12 \sum_{L,M} \obk{w_{\phi-}|D_{LM}} D_{LM}.
+  = \cos\p{qa} \f12 \sum_{L,M} \obk{\tilde d_{\phi+}|D_{LM}} D_{LM}
+  + \sin\p{qa} \f12 \sum_{L,M} \obk{\tilde d_{\phi-}|D_{LM}} D_{LM}.
 \end{align}
@@
 \begin{align}
-  \sum_{L,M} \obk{w_{\phi+}|D_{LM}} D_{LM}
-  = \sum_{L\,\t{even}} \obk{w_{\phi+}|D_{L,0}} D_{L,0},
+  \sum_{L,M} \obk{\tilde d_{\phi+}|D_{LM}} D_{LM}
+  = \sum_{L\,\t{even}} \obk{\tilde d_{\phi+}|D_{L,0}} D_{L,0},
 \end{align}
-and likewise with $w_{\phi-}$ and odd $L$.  In total, we can
+and likewise with $\tilde d_{\phi-}$ and odd $L$.  In total, we can
 write
@@
   = \cos\p{qa} \f12 \sum_{L\,\t{even}}
-  \obk{w_{\phi+}|D_{L,0}} D_{L,0}
+  \obk{\tilde d_{\phi+}|D_{L,0}} D_{L,0}
   + \sin\p{qa} \f12 \sum_{L\,\t{odd}}
-  \obk{w_{\phi-}|D_{L,0}} D_{L,0},
+  \obk{\tilde d_{\phi-}|D_{L,0}} D_{L,0},
 \end{align}
@@
   \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
-  = \sum_L f_L\p{qa} W_L^{(\phi)} D_{L,0},
+  = \sum_L w_L\p{qa} A_L^{(\phi)} D_{L,0},
   &&
-  f_L\p{\theta} \equiv
+  w_L\p{\theta} \equiv
   \begin{cases}
@@
 \begin{align}
-  W_L^{(\phi)}
-  \equiv \f12 \obk{w_{\phi,\p{-1}^L}|D_{L,0}}
-  = \f12 \sum_\mu f_L\p{\mu\phi} \obk{s_{\mu\mu}|D_{L,0}}
-  = \sqrt{\f{L+1/2}{n}}
-  \sum_\mu \bk{I\mu,L,0;I\mu} f_L\p{\mu\phi}.
+  A_L^{(\phi)}
+  \equiv \f12 \obk{\tilde d_{\phi,\p{-1}^L}|D_{L,0}}
+  = \f12 \sum_\mu w_L\p{\mu\phi} \obk{s_{\mu\mu}|D_{L,0}}
+  = \sqrt{\f{2L+1}{n}}
+  \sum_\mu \bk{I\mu;L,0|I\mu} w_L\p{\mu\phi}.
 \end{align}
@@
   &&
-  B_L^{(\phi,q)} \equiv -J f_L\p{qa} W_L^{(\phi)}.
+  B_L^{(\phi,q)} \equiv -J w_L\p{qa} A_L^{(\phi)}.
 \end{align}
@@
 
-Here we compute, for an $n$-level spin with total spin
-$I\equiv\p{n-1}/2$, the structure constants of the transition operator
-algebra, expanding
+Here we derive the structure constants
+$f_{\ell_1 m_1;\ell_2 m_2}^{LM}$ of the transition operators on an
+$n$-dimensional Hilbert space, which allows us to expand
 \begin{align}
-  T_{\ell_2 m_2} T_{\ell_1 m_1}
-  = \sum_{L,M} T_{LM} \obk{T_{LM} | T_{\ell_2 m_2} T_{\ell_1 m_1}},
+  T_{\ell_1 m_1} T_{\ell_2 m_2}
+  = \sum_{L,M} f_{\ell_1 m_1;\ell_2 m_2}^{LM} T_{LM},
 \end{align}
-where the transition operators are defined by
+with
 \begin{align}
-  T_{LM} \equiv \sqrt{\f{2L+1}{2I+1}}
-  \sum_{\mu,\nu} \bk{I\mu,LM;I\nu} \op{\nu}{\mu},
+  f_{\ell_1 m_1;\ell_2 m_2}^{LM}
+  \equiv \obk{T_{LM} | T_{\ell_1 m_1} T_{\ell_2 m_2}}
+  = \tr\p{T_{LM}^\dag T_{\ell_1 m_1} T_{\ell_2 m_2}}.
+  \label{eq:trans_struct}
 \end{align}
-with Clebsch-Gordan coefficients $\bk{I\mu,LM;I\nu}$, and the inner
-product is
+For reference, the transition operators $T_{LM}$ are defined for
+integers $L,M$ with $0\le L<n$ and $\abs{M}\le L$ in terms of
+Clebsch-Gordan coefficients $\bk{\ell_1 m_1;\ell_2 m_2|\ell_3 m_3}$ by
 \begin{align}
-  \obk{T_{LM} | T_{\ell_2 m_2} T_{\ell_1 m_1}}
-  = \tr\p{T_{LM}^\dag T_{\ell_2 m_2} T_{\ell_1 m_1}}
-  = \p{-1}^M \tr\p{T_{L,-M} T_{\ell_2 m_2} T_{\ell_1 m_1}}.
+  T_{LM} \equiv \p{-1}^M \sqrt{\f{2L+1}{2I+1}}
+  \sum_{\mu,\nu} \bk{I\mu;LM|I\nu} \op{\nu}{\mu},
+  &&
+  I \equiv \f{n-1}{2}.
 \end{align}
-The Clebsch-Gordan coefficients can be written in terms of Wigner
-3-$j$ symbols as
+Using the symmetry properties of Clebsch-Gordan coefficients, namely
 \begin{align}
-  \bk{\ell_1 m_1, \ell_2 m_2;\ell_3 m_3}
+  \bk{\ell_1 m_1; \ell_2 m_2| \ell_3 m_3}
+  &= \p{-1}^{\ell_2+m_2} \sqrt{\f{2\ell_3+1}{2\ell_1+1}}
+  \bk{\ell_3,-m_3; \ell_2 m_2| \ell_1,-m_1} \\
+  \bk{\ell_1 m_1; \ell_2 m_2| \ell_3 m_3}
+  &= \p{-1}^{\ell_1+\ell_2-\ell_3}
+  \bk{\ell_1,-m_1; \ell_2,-m_2| \ell_3,-m_3},
+\end{align}
+we can find that
+\begin{align}
+  T_{LM}^\dag
+  = \p{-1}^M \sqrt{\f{2L+1}{2I+1}}
+  \sum_{\mu,\nu} \p{-1}^M \bk{I\nu;L,-M|I\mu} \op{\mu}{\nu}
+  = \p{-1}^M T_{L,-M}.
+\end{align}
+Substituting this result into \eqref{eq:trans_struct} and expanding
+Clebsch-Gordan coefficients in terms of Wigner 3-$j$ symbols as
+\begin{align}
+  \bk{\ell_1 m_1; \ell_2 m_2| \ell_3 m_3}
   = \p{-1}^{-\ell_1+\ell_2-m_3} \sqrt{\p{2\ell_3+1}}
@@
     m_1 & m_2 & -m_3
-  \end{pmatrix}.
+  \end{pmatrix},
 \end{align}
-Substituting transition operators and replacing Clebsch-Gordan
-coefficients by Wigner 3-$j$ symbols, we find that
+we find that
 \begin{multline}
-  \obk{T_{LM} | T_{\ell_2 m_2} T_{\ell_1 m_1}}
-  = \delta_{m_1+m_2,M}
-  \p{-1}^M \sqrt{\p{2\ell_1+1}\p{2\ell_2+1}\p{2L+1}} \\
-  \times \sum_{\mu,\nu,\rho} \p{-1}^{-3I+\ell_1+\ell_2+L-\mu-\nu-\rho}
+  f_{\ell_1 m_1;\ell_2 m_2}^{LM}
+  = \delta_{M,m_1+m_2}
+  \p{-1}^{m_1+m_2} \sqrt{\p{2L+1}\p{2\ell_1+1}\p{2\ell_2+1}} \\
+  \times \sum_{\mu,\nu,\rho} \p{-1}^{-3I+L+\ell_1+\ell_2-\mu-\nu-\rho}
+  \begin{pmatrix}
+    I & L & I \\
+    \mu & -M & -\nu
+  \end{pmatrix}
   \begin{pmatrix}
     I & \ell_1 & I \\
-    \mu & m_1 & -\nu
+    \nu & m_1 & -\rho
   \end{pmatrix}
@@
     I & \ell_2 & I \\
-    \nu & m_2 & -\rho
-  \end{pmatrix}
-  \begin{pmatrix}
-    I & L & I \\
-    \rho & -M & -\mu
+    \rho & m_2 & -\mu
   \end{pmatrix},
@@
 get
-\begin{multline}
-  \obk{T_{LM} | T_{\ell_2 m_2} T_{\ell_1 m_1}} \\
-  = \delta_{m_1+m_2,M} \p{-1}^{2I+M}
-  \sqrt{\p{2\ell_1+1}\p{2\ell_2+1}\p{2L+1}}
+\begin{align}
+  f_{\ell_1 m_1;\ell_2 m_2}^{LM}
+  &= \delta_{M,m_1+m_2} \p{-1}^{2I+M}
+  \sqrt{\p{2L+1}\p{2\ell_1+1}\p{2\ell_2+1}}
   \begin{pmatrix}
-    \ell_1 & \ell_2 & L \\
-    m_1 & m_2 & -M
+    L & \ell_1 & \ell_2 \\
+    M & -m_1 & -m_2
   \end{pmatrix}
   \begin{Bmatrix}
-    \ell_1 & \ell_2 & L \\
-    J & J & J
+    L & \ell_1 & \ell_2 \\
+    I & I & I
   \end{Bmatrix} \\
-  = \delta_{m_1+m_2,M} \p{-1}^{2I+\ell_1-\ell_2}
+  &= \delta_{M,m_1+m_2} \p{-1}^{2I+L}
   \sqrt{\p{2\ell_1+1}\p{2\ell_2+1}}
-  \bk{\ell_1 m_2, \ell_2 m_2; LM}
+  \bk{\ell_1 m_1; \ell_2 m_2| LM}
   \begin{Bmatrix}
     \ell_1 & \ell_2 & L \\
-    J & J & J
+    I & I & I
   \end{Bmatrix}.
+  \label{eq:trans_prod}
+\end{align}
+The Wigner 6-$j$ symbol is symmetric under any permutation of its
+columns, while Clebsch-Gordan coefficients satisfy
+\begin{align}
+  \bk{\ell_1 m_1; \ell_2 m_2| \ell_3 m_3}
+  = \p{-1}^{\ell_1+\ell_2-\ell_3}
+  \bk{\ell_2 m_2; \ell_1 m_1| \ell_3 m_3},
+\end{align}
+which implies that
+\begin{align}
+  f_{\ell_1 m_1;\ell_2 m_2}^{LM}
+  = \p{-1}^{\ell_1+\ell_2-L} f_{\ell_2 m_2;\ell_1 m_1}^{LM}.
+\end{align}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Drive operator product expansions}
+\label{sec:drive_prod}
+
+The drive operators are defined by
+\begin{align}
+  D_{L,0} \equiv T_{L,0},
+  &&
+  D_{LM} \stackrel{M\ne0}{\equiv} \f{\sigma_M}{\sqrt{2}}
+  \sp{T_{L\abs{M}} + \sign\p{M}T_{L\abs{M}}^\dag},
+  &&
+  \sigma_M \equiv \sign\p{M}^{3/2},
+\end{align}
+and can alternately be expressed as
+\begin{align}
+  D_{LM} = \f{\tau_M}{\sqrt{2}} \p{T_{LM} + \epsilon_M T_{L,-M}},
+\end{align}
+where
+\begin{align}
+  \tau_M \equiv
+  \begin{cases}
+    \sqrt{2} & M = 0, \\
+    \sign\p{M}^{M+1/2} & M \ne 0
+  \end{cases},
+  &&
+  \epsilon_M \equiv \p{-1}^M \sign\p{M},
+\end{align}
+which also implies that
+\begin{align}
+  T_{LM} = \f{\tilde\tau_M}{\sqrt{2}} \p{D_{LM} + i D_{L,-M}},
+  &&
+  \tilde\tau_M \equiv
+  \begin{cases}
+    e^{-i\pi/4} & M = 0 \\
+    \sign\p{M}^{M-1/2} & M \ne 0
+  \end{cases}.
+\end{align}
+In order to evaluate the product of two drive operators, we can thus
+expand them in terms of transition operators, evaluate the
+corresponding products of transition operators (see Appendix
+\ref{sec:trans_prod}), and finally expand the result in terms of drive
+operators.  The general product of two drive operators takes the form
+\begin{align}
+  D_{\ell_1 m_1} D_{\ell_2 m_2}
+  = \f{\tau_{m_1}\tau_{m_2}}{\sqrt{2}}
+  \sum_L \p{f_{\ell_1 m_1;\ell_2 m_2}^{L,m_1+m_2}
+    E_{\ell_1 m_1;\ell_2 m_2}^{(L,+)}
+    + f_{\ell_1 m_1;\ell_2,-m_2}^{L,m_1-m_2}
+    E_{\ell_1 m_1;\ell_2 m_2}^{(L,-)}},
+\end{align}
+with operators
+\begin{align}
+  E_{\ell_1 m_1;\ell_2 m_2}^{(L,+)}
+  &\equiv \f12 \sum_{s\in\set{\pm1}} \sqrt{s} D_{L,s\p{m_1+m_2}}
+  \sp{\tilde\tau_{m_1+m_2}
+    + i s \p{-1}^{L+\ell_1+\ell_2}
+    \epsilon_{m_1} \epsilon_{m_2} \tilde\tau_{-\p{m_1+m_2}}},
+  \\
+  E_{\ell_1 m_1;\ell_2 m_2}^{(L,-)}
+  &\equiv \f12 \sum_{s\in\set{\pm1}} \sqrt{s} D_{L,s\p{m_1-m_2}}
+  \sp{\epsilon_{m_2} \tilde\tau_{m_1-m_2}
+    + i s \p{-1}^{L+\ell_1+\ell_2}
+    \epsilon_{m_1} \tilde\tau_{-\p{m_1-m_2}}}.
+\end{align}
+The general product of drive operators simplifies somewhat in several
+special cases.  If one of $m_1,m_2$ is zero, for example, then
+\begin{align}
+  D_{\ell_1 m_1} D_{\ell_2 m_2}
+  \stackrel{0\in\set{m_1,m_2}}{=}
+  \sum_L f_{\ell_1 m_1;\ell_2 m_2}^{L,m_1+m_2} \times
+  \begin{cases}
+    D_{L,m_1+m_2} & L + \ell_1 + \ell_2 ~\t{even} \\
+    i D_{L,-\p{m_1+m_2}} & \t{otherwise}
+  \end{cases}.
+\end{align}
+If $\abs{m_1}=\abs{m_2}$, then
+\begin{multline}
+  D_{\ell_1 m_1} D_{\ell_2 m_2}
+  \stackrel{\abs{m_1}=\abs{m_2}}{=}
+  \sqrt{s_m} \p{-1}^m \sum_L f_{\ell_1 m;\ell_2,-m}^{L,0} D_{L,0} \times
+  \begin{cases}
+    1 & \p{-1}^L = s_m \p{-1}^{\ell_1+\ell_2} \\
+    0 & \t{otherwise}
+  \end{cases} \\
+  + \f{\sqrt{\sign\p{m}}}{\sqrt{s_m}} \f1{\sqrt{2}}
+  \sum_L f_{\ell_1 m;\ell_2 m}^{L,2m} \times
+  \begin{cases}
+    D_{L,2m} & \p{-1}^L = s_m \p{-1}^{\ell_1+\ell_2} \sign\p{m} \\
+    i D_{L,-2m} & \t{otherwise}
+  \end{cases},
 \end{multline}
+where $m \equiv m_1$ and $s_m\equiv+1$ ($-1$) if $m_1=m_2$
+($m_1\ne m_2$).  Finally, if $\abs{m_1}\ne\abs{m_2}$ with
+$m_1,m_2\ne0$, then
+\begin{multline}
+  D_{\ell_1 m_1} D_{\ell_2 m_2}
+  \stackrel{\substack{\abs{m_1}\ne\abs{m_2}\\m_1,m_2\ne0}}{=}
+  \f1{\sqrt{2}} \sum_{s\in\set{\pm1}}\f1{\sqrt{s}}
+  \tau_{m_1} \tau_{s m_2} \tau_{m_1+s m_2}
+  \sign\p{m_1+sm_2} \\
+  \times \sum_L f_{\ell_1 m_1;\ell_2 sm_2}^{L,m_1+sm_2} \times
+  \begin{cases}
+    D_{L,m_1+sm_2} & \p{-1}^L = \zeta_{\ell_1 m_1;\ell_2 m_2}^{(s)} \\
+    i D_{L,-\p{m_1+sm_2}} & \t{otherwise}
+  \end{cases},
+\end{multline}
+where
+\begin{align}
+  \zeta_{\ell_1 m_1;\ell_2 m_2}^{(s)}
+  \equiv s \p{-1}^{\ell_1+\ell_2}
+  \sign\p{m_1} \sign\p{sm_2} \sign\p{m_1+sm_2}.
+\end{align}
 
@@
 considered in Section \ref{sec:drive_raman}.  These coefficients are
-computed by evaluating products of transition operators in
-\eqref{eq:raman_0}--\eqref{eq:raman_2}, using both the structure
-constants of the transition operator algebra (derived in Appendix
-\ref{sec:trans_prod}) and the fact that
-$T_{LM}^\dag = \p{-1}^M T_{L,-M}$.  For clarity, if any coefficient
-$\Omega_{LM}^{(\phi,j)}$ is independent of $\phi$, we drop the
-explicit dependence on $\phi$, i.e.~taking
+computed using symmetry properties and structure constants of the
+transition operator algebra (see Appendix \ref{sec:trans_prod}).  For
+clarity, if any coefficient $\Omega_{LM}^{(\phi,j)}$ is independent of
+$\phi$, we drop the explicit dependence on $\phi$, i.e.~taking
 $\Omega_{LM}^{(\phi,j)}\to\Omega_{LM}$.
 \begin{align}
-  \Omega_{0,0} \equiv -\f1{\sqrt{2}}~ \f1{\sqrt{n}}~
+  \Omega_{0,0} \equiv -\f1{\sqrt{n}}\,
   \sum_m\f{\Omega_m^2}{\Delta},
   &&
-  \Omega_{1,0} \equiv \sqrt{\f32}~ \f1{\sqrt{n_1}}~
+  \Omega_{1,0} \equiv \sqrt{\f{3}{n_1}}\,
   \f{\Omega_+^2 - \Omega_-^2}{\Delta},
@@
 \begin{align}
-  \Omega_{2,0} \equiv \f1{\sqrt{10}}~ \f{\sqrt{n_2}}{n_1}~
+  \Omega_{2,0} \equiv \f1{\sqrt{5}}\, \f{\sqrt{n_2}}{n_1}\,
   \f{\Omega_+ + \Omega_- - 2\Omega_0}{\Delta},
@@
   \omega_{1\pm}
-  &\equiv \sqrt{3} ~\f1{\sqrt{n_1}}~
-  \f{\Omega_0 \Omega_\pm}{\Delta},
+  &\equiv -\sqrt{\f{6}{n_1}}\, \f{\Omega_0 \Omega_\pm}{\Delta},
   &
   \omega_{2\pm}
-  &\equiv \pm \sqrt{\f35}~ \f{\sqrt{n_2}}{n_1}~
+  &\equiv \mp \sqrt{\f65}\, \f{\sqrt{n_2}}{n_1}\,
   \f{\Omega_0\Omega_\pm}{\Delta}, \\
   \omega_{2,2}
-  &\equiv -\sqrt{\f65}~ \f{\sqrt{n_2}}{n_1}~
+  &\equiv 2\sqrt{\f35}\, \f{\sqrt{n_2}}{n_1}\,
   \f{\Omega_+\Omega_-}{\Delta},
@@
   n_L
-  &\equiv \prod_{\ell=-L}^L \p{n-\ell}.
+  &\equiv \prod_{\ell=-L}^L \p{n+\ell}.
 \end{align}
@@
 $\phi_-=0$, or equivalently re-define $\phi_m-\phi_0\to\phi_m$.  If
-the driving lasers are phase-matched with $\phi_s=s\phi$, such that a
-particle on site $j$ acquires a phase $\mp\phi j$ for every nuclear
-spin transition $\mu\to\mu\pm1$, then the drive
-$H_{\t{Raman}}^{(\phi)}$ takes the form
+the driving lasers are phase-matched with $\phi_m=m\phi$, such that a
+nuclear spin transition $\mu\to\mu+m$ on site $j$ imprints the phase
+$-m\phi j$, then the drive $H_{\t{Raman}}^{(\phi)}$ takes the form
 \begin{align}
-  \left. H_{\t{Raman}}^{(\phi)} \right|_{\phi_s=s\phi}
+  \left. H_{\t{Raman}}^{(\phi)} \right|_{\phi_m=m\phi}
   = \sum_{\substack{L\le2\\0\le M\le L}} \Omega_{LM} \D_{LM}^{(\phi)},

### 2019-10-10 collective_sun/sun_notes.tex
simplified work in drive operator product appendix
removing item from to-do list
fixed error in drive operator product appendix
minor simplifications and additions to drive operator product section

@@
 \begin{itemize}
-\item Products and commutators of drive operators
 \item Visualization with generalized Bloch spheres
@@
   \begin{cases}
-    e^{-i\pi/4} & M = 0 \\
+    \sqrt{2}/\p{1+i} & M = 0 \\
     \sign\p{M}^{M-1/2} & M \ne 0
@@
   E_{\ell_1 m_1;\ell_2 m_2}^{(L,+)}
-  &\equiv \f12 \sum_{s\in\set{\pm1}} \sqrt{s} D_{L,s\p{m_1+m_2}}
-  \sp{\tilde\tau_{m_1+m_2}
+  &\equiv \sum_{s\in\set{\pm1}}
+  \f12 \sp{\tilde\tau_{m_1+m_2}
     + i s \p{-1}^{L+\ell_1+\ell_2}
-    \epsilon_{m_1} \epsilon_{m_2} \tilde\tau_{-\p{m_1+m_2}}},
+    \epsilon_{m_1} \epsilon_{m_2} \tilde\tau_{-\p{m_1+m_2}}}
+  \sqrt{s} D_{L,s\p{m_1+m_2}},
   \\
   E_{\ell_1 m_1;\ell_2 m_2}^{(L,-)}
-  &\equiv \f12 \sum_{s\in\set{\pm1}} \sqrt{s} D_{L,s\p{m_1-m_2}}
-  \sp{\epsilon_{m_2} \tilde\tau_{m_1-m_2}
+  &\equiv \sum_{s\in\set{\pm1}}
+  \f12 \sp{\epsilon_{m_2} \tilde\tau_{m_1-m_2}
     + i s \p{-1}^{L+\ell_1+\ell_2}
-    \epsilon_{m_1} \tilde\tau_{-\p{m_1-m_2}}}.
+    \epsilon_{m_1} \tilde\tau_{-\p{m_1-m_2}}}
+  \sqrt{s} D_{L,s\p{m_1-m_2}}.
 \end{align}
-The general product of drive operators simplifies somewhat in several
-special cases.  If one of $m_1,m_2$ is zero, for example, then
+If $m_1\pm m_2=0$, these operators simplify to
 \begin{align}
-  D_{\ell_1 m_1} D_{\ell_2 m_2}
-  \stackrel{0\in\set{m_1,m_2}}{=}
-  \sum_L f_{\ell_1 m_1;\ell_2 m_2}^{L,m_1+m_2} \times
-  \begin{cases}
-    D_{L,m_1+m_2} & L + \ell_1 + \ell_2 ~\t{even} \\
-    i D_{L,-\p{m_1+m_2}} & \t{otherwise}
-  \end{cases}.
+  E_{\ell_1 m;\ell_2,-m}^{(L,+)}
+  &= \f1{\sqrt{2}} \sp{1 - \p{-1}^{L+\ell_1+\ell_2}
+    \p{1-\delta_{m,0}}} D_{L,0}, \\
+  E_{\ell_1 m;\ell_2 m}^{(L,-)}
+  &= \f{\epsilon_m}{\sqrt{2}}
+  \sp{1 + \p{-1}^{L+\ell_1+\ell_2}} D_{L,0},
 \end{align}
-If $\abs{m_1}=\abs{m_2}$, then
-\begin{multline}
-  D_{\ell_1 m_1} D_{\ell_2 m_2}
-  \stackrel{\abs{m_1}=\abs{m_2}}{=}
-  \sqrt{s_m} \p{-1}^m \sum_L f_{\ell_1 m;\ell_2,-m}^{L,0} D_{L,0} \times
-  \begin{cases}
-    1 & \p{-1}^L = s_m \p{-1}^{\ell_1+\ell_2} \\
-    0 & \t{otherwise}
-  \end{cases} \\
-  + \f{\sqrt{\sign\p{m}}}{\sqrt{s_m}} \f1{\sqrt{2}}
-  \sum_L f_{\ell_1 m;\ell_2 m}^{L,2m} \times
-  \begin{cases}
-    D_{L,2m} & \p{-1}^L = s_m \p{-1}^{\ell_1+\ell_2} \sign\p{m} \\
-    i D_{L,-2m} & \t{otherwise}
-  \end{cases},
-\end{multline}
-where $m \equiv m_1$ and $s_m\equiv+1$ ($-1$) if $m_1=m_2$
-($m_1\ne m_2$).  Finally, if $\abs{m_1}\ne\abs{m_2}$ with
-$m_1,m_2\ne0$, then
-\begin{multline}
-  D_{\ell_1 m_1} D_{\ell_2 m_2}
-  \stackrel{\substack{\abs{m_1}\ne\abs{m_2}\\m_1,m_2\ne0}}{=}
-  \f1{\sqrt{2}} \sum_{s\in\set{\pm1}}\f1{\sqrt{s}}
-  \tau_{m_1} \tau_{s m_2} \tau_{m_1+s m_2}
-  \sign\p{m_1+sm_2} \\
-  \times \sum_L f_{\ell_1 m_1;\ell_2 sm_2}^{L,m_1+sm_2} \times
-  \begin{cases}
-    D_{L,m_1+sm_2} & \p{-1}^L = \zeta_{\ell_1 m_1;\ell_2 m_2}^{(s)} \\
-    i D_{L,-\p{m_1+sm_2}} & \t{otherwise}
-  \end{cases},
-\end{multline}
-where
+and are otherwise
 \begin{align}
-  \zeta_{\ell_1 m_1;\ell_2 m_2}^{(s)}
-  \equiv s \p{-1}^{\ell_1+\ell_2}
-  \sign\p{m_1} \sign\p{sm_2} \sign\p{m_1+sm_2}.
+  E_{\ell_1 m_1;\ell_2 m_2}^{(L,+)}
+  &\stackrel{m_1+m_2\ne0}{=}
+  \tau_{m_1+m_2}^* \sum_{s\in\set{\pm1}}
+  \f12 \sp{1 + s \p{-1}^{L+\ell_1+\ell_2}
+    \epsilon_{m_1} \epsilon_{m_2} \epsilon_{m_1+m_2}}
+  \sqrt{s} D_{L,s\p{m_1+m_2}}, \\
+  E_{\ell_1 m_1;\ell_2 m_2}^{(L,-)}
+  &\stackrel{m_1-m_2\ne0}{=}
+  \tau_{m_1-m_2}^* \sum_{s\in\set{\pm1}}
+  \f12 \sp{\epsilon_{m_2} + s \p{-1}^{L+\ell_1+\ell_2}
+    \epsilon_{m_1} \epsilon_{m_1-m_2}}
+  \sqrt{s} D_{L,s\p{m_1-m_2}}.
+\end{align}
+If $\p{m_1,m_2}=\p{0,m}$, then we can use the fact that
+$i\tau_m\tau_{-m}^*=\epsilon_m$ for $m\ne0$ to simplify
+\begin{align}
+  D_{\ell_1,0} D_{\ell_2 m}
+  = \sum_L f_{\ell_1,0,\ell_2,m}^{Lm}
+  \sum_{s\in\set{\pm1}} \f12 \sp{1 + s \p{-1}^{L+\ell_1+\ell_2}}
+  \sqrt{s} D_{L,sm}.
 \end{align}

### 2019-10-10 collective_sun/sun_notes.tex
changed sign conventions in the definition of the transition operator new sign convention is consistent with existing literature
changed sign conventions in SOC phase

@@
   T_{LM}
-  \equiv \p{-1}^M \sqrt{\f{2L+1}{2I+1}} \sum_{\mu,\nu}
+  \equiv \sqrt{\f{2L+1}{2I+1}} \sum_{\mu,\nu}
   \bk{I\mu;LM|I\nu} \op{\nu}{\mu},
@@
 $n$-level nuclear spin onto a quantization axis; and
-$\bk{I\mu;LM|I\nu}$ is a Clebsch-Gordan coefficient.  Note that our
-definition of a transition operators differs from that in
-Refs.~\cite{kryszewski2006alternative, bertlmann2008bloch} by the sign
-$\p{-1}^M$; this choice of sign convention will simplify much of our
-work below.
-
-The transition operator $T_{LM}$ is proportional to the transition
-induced on a nuclear spin by the absorption of a spin-$L$ particle
-with spin projection $M$ onto a quantization axis.  Similarly to the
-GGM operators in \eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag}, the
+$\bk{I\mu;LM|I\nu}$ is a Clebsch-Gordan coefficient.  The transition
+operator $T_{LM}$ is proportional to the transition induced on a
+nuclear spin by the absorption of a spin-$L$ particle with spin
+projection $M$ onto a quantization axis.  Similarly to the GGM
+operators in \eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag}, the
 transition operators in \eqref{eq:trans_ops} provide a complete basis
@@
   &&
-  \sigma_M \equiv \sign\p{M}^{3/2},
+  \sigma_M \equiv \p{-1}^M \sign\p{M}^{3/2},
   \label{eq:drive_ops}
@@
 
-An driving laser addressing spin-$1/2$ fermions on a 1-D lattice will
+A driving laser addressing spin-$1/2$ fermions on a 1-D lattice will
 induce the Hamiltonian
@@
   = \Omega \sum_j
-  \p{e^{-i\phi j} c_{j,\up}^\dag c_{j,\dn}
-    + e^{i\phi j} c_{j,\dn}^\dag c_{j,\up}},
+  \p{e^{i\phi j} c_{j,\up}^\dag c_{j,\dn}
+    + e^{-i\phi j} c_{j,\dn}^\dag c_{j,\up}},
   \label{eq:H_drive_2}
@@
   H_{\t{drive}}^{(\phi)}
-  = -\Omega \sum_j \p{e^{-i\phi j} T_{1,1}^{(j)}
-    + e^{i\phi j} {T_{1,1}^{(j)}}^\dag}
+  = -\Omega \sum_j \p{e^{i\phi j} T_{1,1}^{(j)}
+    + e^{-i\phi j} {T_{1,1}^{(j)}}^\dag}
   = \Omega \D_{1,1}^{(\phi)},
@@
   \equiv \sum_j \sp{\cos\p{M\phi j} D_{LM}^{(j)}
-    + \sin\p{M\phi j} D_{L,-M}^{(j)}},
+    - \sin\p{M\phi j} D_{L,-M}^{(j)}},
   \label{eq:drive_rot}
@@
 this drive spatially homogenous via the gauge transformation
-$c_{j\mu}^\dag \to e^{i\mu\phi j} c_{j\mu}^\dag$, which takes
+$c_{j\mu}^\dag \to e^{-i\mu\phi j} c_{j\mu}^\dag$, which takes
 \begin{align}
@@
   \to H_{\t{lat}}^{(\phi)}
-  = -J \sum_{q,\mu} \cos\p{qa-\mu\phi} c_{q\mu}^\dag c_{q\mu}
+  = -J \sum_{q,\mu} \cos\p{qa+\mu\phi} c_{q\mu}^\dag c_{q\mu}
   = \sum_{q,L} B_L^{(\phi,q)} D_{L,0}^{(q)},
@@
 \begin{align}
-  c_L \equiv \p{-1}^{\lfloor L/2\rfloor}
+  c_L \equiv \p{-1}^{L+\lfloor L/2\rfloor}
   \sqrt{2L+1}\, \f{L!}{\p{2L+1}!},
@@
   H_{\t{Raman,bare}}^{(\phi)}
-  = \sum_j \sum_{m\in\set{1,0,-1}} \p{\Omega_m e^{-i\phi_m j}
+  = \sum_j \sum_{m\in\set{1,0,-1}} \p{\Omega_m e^{i\phi_m j}
     T_{1,m}^{(j)} \otimes \op{\up}{\dn} + \t{h.c.}}
@@
   &= -\sum_j\sum_{s\in\set{\pm1}}
-  \f{\Omega_0\Omega_s}{\Delta} e^{-i\p{\phi_s-\phi_0} j}
+  \f{\Omega_0\Omega_s}{\Delta} e^{i\p{\phi_s-\phi_0} j}
   {T_{1,0}^{(j)}}^\dag T_{1,s}^{(j)} + \t{h.c.},
@@
   &= -\sum_j\sum_{s\in\set{\pm1}}
-  \f{\Omega_+\Omega_-}{\Delta} e^{-i\p{\phi_s-\phi_{-s}}j}
+  \f{\Omega_+\Omega_-}{\Delta} e^{i\p{\phi_s-\phi_{-s}}j}
   {T_{1,-s}^{(j)}}^\dag T_{1,s}^{(j)}.
@@
 phase-matched with $\phi_m=m\phi$, such that a nuclear spin transition
-$\mu\to\mu+m$ on site $j$ imprints the phase $-m\phi j$, then the
-drive $H_{\t{Raman}}^{(\phi)}$ takes the form
+$\mu\to\mu+m$ on site $j$ imprints the phase $m\phi j$, then the drive
+$H_{\t{Raman}}^{(\phi)}$ takes the form
 \begin{align}
@@
   H_{\t{lat}}^{(\phi)}
-  = -J \sum_{q,\mu} \cos\p{qa-\mu\phi} s_{\mu\mu}^{(q)}
+  = -J \sum_{q,\mu} \cos\p{qa+\mu\phi} s_{\mu\mu}^{(q)}
   \label{eq:H_lat_spin}
@@
   \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
-  = \cos\p{qa} \tilde d_{\phi+} + \sin\p{qa} \tilde d_{\phi-},
+  = \cos\p{qa} \tilde d_{\phi+} - \sin\p{qa} \tilde d_{\phi-},
 \end{align}
@@
   = \cos\p{qa} \f12 \sum_{L,M} \obk{\tilde d_{\phi+}|D_{LM}} D_{LM}
-  + \sin\p{qa} \f12 \sum_{L,M} \obk{\tilde d_{\phi-}|D_{LM}} D_{LM}.
+  - \sin\p{qa} \f12 \sum_{L,M} \obk{\tilde d_{\phi-}|D_{LM}} D_{LM}.
 \end{align}
@@
   \obk{\tilde d_{\phi+}|D_{L,0}} D_{L,0}
-  + \sin\p{qa} \f12 \sum_{L\,\t{odd}}
+  - \sin\p{qa} \f12 \sum_{L\,\t{odd}}
   \obk{\tilde d_{\phi-}|D_{L,0}} D_{L,0},
@@
   A_L^{(\phi)}
-  \equiv \f12 \obk{\tilde d_{\phi,\p{-1}^L}|D_{L,0}}
-  = \f12 \sum_\mu w_L\p{\mu\phi} \obk{s_{\mu\mu}|D_{L,0}}
-  = \sqrt{\f{2L+1}{n}}
+  \equiv \p{-1}^L \f12 \obk{\tilde d_{\phi,\p{-1}^L}|D_{L,0}}
+  = \p{-1}^L \sqrt{\f{2L+1}{n}}
   \sum_\mu \bk{I\mu;L,0|I\mu} w_L\p{\mu\phi}.
@@
   &&
-  \sigma_M \equiv \sign\p{M}^{3/2},
+  \sigma_M \equiv \p{-1}^M \sign\p{M}^{3/2},
 \end{align}
@@
     \sqrt{2} & M = 0, \\
-    \sign\p{M}^{M+1/2} & M \ne 0
+    \p{-1}^M \sign\p{M}^{M+1/2} & M \ne 0
   \end{cases},
@@
     \sqrt{2}/\p{1+i} & M = 0 \\
-    \sign\p{M}^{M-1/2} & M \ne 0
+    \tau_M^* & M \ne 0
   \end{cases}.
@@
   \Omega_{L,-1}^{(\phi,j)}
-  &\equiv \sum_{s\in\set{\pm1}} s \omega_{Ls} \sin\p{\phi_s j}, \\
+  &\equiv -\sum_{s\in\set{\pm1}} s \omega_{Ls} \sin\p{\phi_s j}, \\
   \Omega_{2,2}^{(\phi,j)}
@@
   \Omega_{2,-2}^{(\phi,j)}
-  &\equiv \omega_{2,2} \sin\p{\sp{\phi_+-\phi_-}j}, \\
+  &\equiv -\omega_{2,2} \sin\p{\sp{\phi_+-\phi_-}j}, \\
   \omega_{1\pm}
-  &\equiv -\sqrt{\f{6}{n_1}}\, \f{\Omega_0 \Omega_\pm}{\Delta},
+  &\equiv \sqrt{\f{6}{n_1}}\, \f{\Omega_0 \Omega_\pm}{\Delta},
   &
   \omega_{2\pm}
-  &\equiv \mp \sqrt{\f65}\, \f{\sqrt{n_2}}{n_1}\,
+  &\equiv \pm \sqrt{\f65}\, \f{\sqrt{n_2}}{n_1}\,
   \f{\Omega_0\Omega_\pm}{\Delta}, \\
@@
 nuclear spin transition $\mu\to\mu+m$ on site $j$ imprints the phase
-$-m\phi j$, then the drive $H_{\t{Raman}}^{(\phi)}$ takes the form
+$m\phi j$, then the drive $H_{\t{Raman}}^{(\phi)}$ takes the form
 \begin{align}
@@
   \equiv \sum_j \sp{\cos\p{M\phi j} D_{LM}^{(j)}
-    + \sin\p{M\phi j} D_{L,-M}^{(j)}},
+    - \sin\p{M\phi j} D_{L,-M}^{(j)}},
 \end{align}

### 2019-10-13 collective_sun/multilevel_spin_notes.tex
fixed some minor errors and started section on classifying collective states

--- 2019-10-10 collective_sun/sun_notes.tex
@@
 
+\renewcommand{\SS}{\mathbb{S}}
+
 \DeclareMathOperator{\sign}{sign}
-\DeclareMathOperator{\signs}{signs}
 
@@
 $s_{\mu\nu}^{(p)}$ in terms of Pauli operators in order to convert
-$H_{\t{int}}$ into an SU(2) spin Hamiltonian.  To generalize this
-procedure to the case of SU($n$), we first define a vector
-$\v s^{(p)}$ of all spin operators $s_{\mu\nu}^{(p)}$, and write
+$H_{\t{int}}$ into an SU(2) spin Hamiltonian\cite{he2019engineering}.
+To generalize this procedure to the case of SU($n$), we first define a
+vector $\v s^{(p)}$ of all spin operators $s_{\mu\nu}^{(p)}$, and
+write
 \begin{align}
@@
 $\bk{I\mu;LM|I\nu}$ is a Clebsch-Gordan coefficient.  The transition
-operator $T_{LM}$ is proportional to the transition induced on a
-nuclear spin by the absorption of a spin-$L$ particle with spin
-projection $M$ onto a quantization axis.  Similarly to the GGM
-operators in \eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag}, the
-transition operators in \eqref{eq:trans_ops} provide a complete basis
-for the space of operators on the Hilbert space of an $n$-level
-system, and satisfy the orthonormality condition
+operator $T_{LM}$ is closely related to the associated Legendre
+polynomial $P_{LM}\p{x}$ for $x\in\sp{-1,1}$.  Operationally, $T_{LM}$
+is proportional to the transition induced on a nuclear spin by the
+absorption of a spin-$L$ boson with spin projection $M$ onto a
+quantization axis.
+
+Similarly to the GGM operators in \eqref{eq:GGM_diag} and
+\eqref{eq:GGM_off_diag}, the transition operators in
+\eqref{eq:trans_ops} provide a complete basis for the space of
+operators on the Hilbert space of an $n$-level system, and satisfy the
+orthonormality condition
 \begin{align}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Drive opermators}
+\subsection{Drive operators}
 \label{sec:drive_ops}
@@
 \begin{align}
-  D_{L,0} \equiv T_{L,0},
-  &&
-  D_{LM} \stackrel{M\ne0}{\equiv} \f{\sigma_M}{\sqrt{2}}
+  D_{LM} \equiv \f{\tau_M}{\sqrt{2}}
   \sp{T_{L\abs{M}} + \sign\p{M}T_{L\abs{M}}^\dag},
   &&
-  \sigma_M \equiv \p{-1}^M \sign\p{M}^{3/2},
+  \tau_M \equiv
+  \begin{cases}
+    \sqrt{2} & M = 0 \\
+    \p{-1}^M \sign\p{M}^{3/2} & M \ne 0
+  \end{cases},
   \label{eq:drive_ops}
@@
 one can identify $\p{D_{1,0},D_{1,1},D_{1,-1}}\sim\p{S_\z,S_\x,S_\y}$,
-where $S_\alpha$ is a collective SU(2) spin-$\alpha$
-operator\cite{perlin2019shorttime}.  More concretely, we can embed the
-Hilbert space of a single $n$-level spin into the
+where $S_\alpha$ is a collective SU(2) spin-$\alpha$ operator for
+$\p{n-1}$ 2-level spins\cite{perlin2019shorttime}.  More concretely,
+we can embed the Hilbert space of a single $n$-level spin into the
 permutationally-symmetric (Dicke) manifold of states for $\p{n-1}$
@@
   \times \f{2\sqrt{3}}{\sqrt{\p{n-1}n\p{n+1}}}.
+  \label{eq:spin_ops}
+\end{align}
+Furthermore, if we {\it define} the operators $S_\alpha$ for a single
+$n$-level spin via the correspondence in \eqref{eq:spin_ops}, then the
+familiar SU(2) transformation rule
+\begin{align}
+  e^{i\phi S_\z} S_\x e^{-i\phi S_\z} = S_\x^{(\phi)},
+  &&
+  S_\x^{(\phi)} \equiv \cos\phi S_\x - \sin\phi S_\y,
+\end{align}
+generalizes to
+\begin{align}
+  e^{i\phi S_\z} D_{LM} e^{-i\phi S_\z} = D_{LM}^{(\phi)},
+  &&
+  D_{LM}^{(\phi)} \equiv \cos\p{M\phi} D_{LM} - \sin\p{M\phi} D_{L,-M}.
 \end{align}
 
-We note that Weyl operators\cite{bertlmann2008bloch} and self-adjoint
-linear combinations thereof\cite{asadian2016heisenbergweyl} also
-provide an interesting basis of operators, generalizing continuous
-phase-space displacement operators to the case of a finite-dimensional
-Hilbert space.  As our present work is motivated by multi-level
-systems that are realized with nuclear spin (angular momentum) degrees
-of freedom, however, the transition and drive operators are more
-natural to consider than the Weyl operators, so we will not discuss
-the latter any further.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Weyl operators}
+
+Weyl operators\cite{bertlmann2008bloch} and self-adjoint linear
+combinations thereof\cite{asadian2016heisenbergweyl} also provide an
+interesting basis of operators for finite-dimensional Hilbert spaces.
+These operators are related to the notions of position, momentum, and
+phase-space displacement in a discrete space.  As our present work is
+motivated by multi-level systems that are realized with nuclear spin
+(angular momentum) degrees of freedom, however, the transition and
+drive operators are more natural to use than the Weyl operators, so we
+will not discuss Weyl operators any further.
 
@@
   &&
-  A_L^{(\phi)} \equiv \sqrt{\f{2L+1}{n}}
+  A_L^{(\phi)} \equiv \p{-1}^L \sqrt{\f{2L+1}{2I+1}}
   \sum_\mu \bk{I\mu;L,0|I\mu} w_L\p{\mu\phi}.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Classifying collective states}
+
+Given a quantum system composed of $N$ identical (but distinguishable)
+sub-systems, e.g.~an array of $n$-level nuclear spins, we identify the
+set $\Pi\equiv\set{\Pi_\sigma}$ of operators that permute sub-systems,
+labeled by elements $\sigma\in\SS_N$ of the symmetric group $\SS_N$ of
+order $N$.  We refer to simultaneous eigenstates of all
+$\Pi_\sigma\in\Pi$ as {\it collective states}.  Collective states that
+have eigenvalue 1 with respect to all $\Pi_\sigma\in\Pi$ generalize
+the notion of Dicke states to the case of multi-level sub-systems.
+
+\vspace{3cm}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Future directions}
@@
   \begin{itemize}
-  \item Effective collective dynamics for $\abs{\phi}\ll1$
-  \item Deriving \eqref{eq:A_L_phi_small}
   \item Special angles $\phi$
@@
   \equiv \p{-1}^L \f12 \obk{\tilde d_{\phi,\p{-1}^L}|D_{L,0}}
-  = \p{-1}^L \sqrt{\f{2L+1}{n}}
+  = \p{-1}^L \sqrt{\f{2L+1}{2I+1}}
   \sum_\mu \bk{I\mu;L,0|I\mu} w_L\p{\mu\phi}.
@@
 \begin{align}
-  D_{L,0} \equiv T_{L,0},
-  &&
-  D_{LM} \stackrel{M\ne0}{\equiv} \f{\sigma_M}{\sqrt{2}}
+  D_{LM} \equiv \f{\tau_M}{\sqrt{2}}
   \sp{T_{L\abs{M}} + \sign\p{M}T_{L\abs{M}}^\dag},
   &&
-  \sigma_M \equiv \p{-1}^M \sign\p{M}^{3/2},
-\end{align}
-and can alternately be expressed as
-\begin{align}
-  D_{LM} = \f{\tau_M}{\sqrt{2}} \p{T_{LM} + \epsilon_M T_{L,-M}},
-\end{align}
-where
-\begin{align}
   \tau_M \equiv
   \begin{cases}
-    \sqrt{2} & M = 0, \\
-    \p{-1}^M \sign\p{M}^{M+1/2} & M \ne 0
+    \sqrt{2} & M = 0 \\
+    \p{-1}^M \sign\p{M}^{3/2} & M \ne 0
   \end{cases},
-  &&
-  \epsilon_M \equiv \p{-1}^M \sign\p{M},
 \end{align}
-which also implies that
+which implies that
 \begin{align}
@@
 \end{align}
-In order to evaluate the product of two drive operators, we can thus
-expand them in terms of transition operators, evaluate the
-corresponding products of transition operators (see Appendix
-\ref{sec:trans_prod}), and finally expand the result in terms of drive
-operators.  The general product of two drive operators takes the form
+where $\tau_M^*$ is the complex conjugate of $\tau_M$.  In order to
+evaluate the product of two drive operators, we can thus expand them
+in terms of transition operators, evaluate the corresponding products
+of transition operators (see Appendix \ref{sec:trans_prod}), and
+finally expand the result in terms of drive operators.  The general
+product of two drive operators takes the form
 \begin{align}
@@
 
-\bibliography{sun_notes.bib}
+\bibliography{multilevel_spin_notes.bib}
 

### 2019-10-13 collective_sun/multilevel_spin_notes.tex
fixed minor error from old change in sign conventions for transition operators

@@
 have eigenvalue 1 with respect to all $\Pi_\sigma\in\Pi$ generalize
-the notion of Dicke states to the case of multi-level sub-systems.
+Dicke states to the case of multi-level sub-systems.
 
@@
 
-Here we derive the structure constants
-$f_{\ell_1 m_1;\ell_2 m_2}^{LM}$ of the transition operators on an
-$n$-dimensional Hilbert space, which allows us to expand
+Here we derive multiplication rules for transition operators on an $n$-dimensional Hilbert space, which allows us to expand
 \begin{align}
@@
 \end{align}
-with
+with structure constants
 \begin{align}
@@
 \begin{align}
-  T_{LM} \equiv \p{-1}^M \sqrt{\f{2L+1}{2I+1}}
+  T_{LM} \equiv \sqrt{\f{2L+1}{2I+1}}
   \sum_{\mu,\nu} \bk{I\mu;LM|I\nu} \op{\nu}{\mu},
@@
   T_{LM}^\dag
-  = \p{-1}^M \sqrt{\f{2L+1}{2I+1}}
+  = \sqrt{\f{2L+1}{2I+1}}
   \sum_{\mu,\nu} \p{-1}^M \bk{I\nu;L,-M|I\mu} \op{\mu}{\nu}
@@
   = \delta_{M,m_1+m_2}
-  \p{-1}^{m_1+m_2} \sqrt{\p{2L+1}\p{2\ell_1+1}\p{2\ell_2+1}} \\
+  \p{-1}^M \sqrt{\p{2L+1}\p{2\ell_1+1}\p{2\ell_2+1}} \\
   \times \sum_{\mu,\nu,\rho} \p{-1}^{-3I+L+\ell_1+\ell_2-\mu-\nu-\rho}

### 2019-10-14 collective_sun/multilevel_spin_notes.tex
added proof that the maximally-symmetric manifold is closed under the action of unitaries generated by collective operators

@@
 \newcommand{\D}{\mathcal{D}}
+\renewcommand{\H}{\mathcal{H}}
 \newcommand{\I}{\mathcal{I}}
@@
 \renewcommand{\S}{\mathcal{S}}
+\newcommand{\X}{\mathcal{X}}
 
 \renewcommand{\SS}{\mathbb{S}}
+\newcommand{\ZZ}{\mathbb{Z}}
 
@@
 internal state (e.g.~nuclear spin) $\mu$.  Under the frozen-mode
-approximation, two-body collective SU($n$)-symmetric interactions
+approximation, collective two-body SU($n$)-symmetric interactions
 between these fermions take form\cite{perlin2019effective}
@@
 In the remainder of this section, we identify various bases for the
-space of operators on the Hilbert space of an $n$-level system.  For
-brevity, we denote this space of operators by $\B_n$, and note that
-$\B_n$ is itself an $n^2$-dimensional Hilbert space equipped with a
-trace (Hilbert-Schmidt) inner product
+space of linear operators on the Hilbert space $\H_n$ of an $n$-level
+system.  For brevity, we denote this space of operators by
+$\B\p{\H_n}$, and note that $\B\p{\H_n}$ is itself an
+$n^2$-dimensional Hilbert space equipped with a trace
+(Hilbert-Schmidt) inner product
 \begin{align}
@@
 where $\oket{\Q}$ and $\obra{\O}$ respectively denote vectors in
-$\B_n$ and its dual space $\B_n^*$.  Similarly to the set of all spin
-operators $s_{\mu\nu}\equiv\op{\mu}{\nu}$ on the single-particle
-Hilbert space at each quasi-momentum, all of the bases we consider
-will be orthonormal with respect to the trace inner product.  As we
-show in Appendix \ref{sec:changing_bases}, the collective
-SU($n$)-symmetric interaction Hamiltonian in \eqref{eq:H_int_spin}
-takes an identical form in any orthonormal basis for $\B_n$.
+$\B\p{\H_n}$ and its dual space $\B\p{\H_n}^*$.  Similarly to the set
+of all spin operators $s_{\mu\nu}\equiv\op{\mu}{\nu}$ on the
+single-particle Hilbert space at each quasi-momentum, all of the bases
+we consider will be orthonormal with respect to the trace inner
+product.  As we show in Appendix \ref{sec:changing_bases}, the
+collective SU($n$)-symmetric interaction Hamiltonian in
+\eqref{eq:H_int_spin} takes an identical form in any orthonormal basis
+for $\B\p{\H_n}$.
 
@@
 in \eqref{eq:GGM_diag} and \eqref{eq:GGM_off_diag} provide a complete
-basis for the space of operators on the Hilbert space of an $n$-level
-system, coincide exactly with the Pauli operators in the case of
-$n=2$, and satisfy the orthonormality condition
+basis for the space of operators on the Hilbert space $\H_n$ of an
+$n$-level system, coincide exactly with the Pauli operators in the
+case of $n=2$, and satisfy the orthonormality condition
 \begin{align}
@@
 \eqref{eq:trans_ops} provide a complete basis for the space of
-operators on the Hilbert space of an $n$-level system, and satisfy the
-orthonormality condition
+operators on the Hilbert space $\H_n$ of an $n$-level system, and
+satisfy the orthonormality condition
 \begin{align}
@@
  \begin{align}
-  H_{\t{drive}}^{(\phi)} \to H_{\t{drive}} = \Omega \D_{1,1}.
+  H_{\t{drive}}^{(\phi)} \to H_{\t{drive}} = \Omega \D_{1,1},
+  &&
+  \D_{LM} \equiv \D_{LM}^{(0)} = \sum_j D_{LM}^{(j)}.
 \end{align}
@@
 \end{enumerate*}
-(and conjecturing the same pattern for all $L$ and $n$), we find that
+(and conjecturing the same functional form for all $L$ and $n$), we
+find that
 % expand Clebsch-Gordan coefficients in a power series according to
@@
 \section{Classifying collective states}
+\label{sec:classifying}
 
 Given a quantum system composed of $N$ identical (but distinguishable)
-sub-systems, e.g.~an array of $n$-level nuclear spins, we identify the
-set $\Pi\equiv\set{\Pi_\sigma}$ of operators that permute sub-systems,
-labeled by elements $\sigma\in\SS_N$ of the symmetric group $\SS_N$ of
-order $N$.  We refer to simultaneous eigenstates of all
-$\Pi_\sigma\in\Pi$ as {\it collective states}.  Collective states that
-have eigenvalue 1 with respect to all $\Pi_\sigma\in\Pi$ generalize
-Dicke states to the case of multi-level sub-systems.
+subsystems, e.g.~an array of $n$-level nuclear spins, we identify the
+set $\Pi\equiv\set{\Pi_\sigma}$ of operators that permute subsystems.
+Permutation operators $\Pi_\sigma$ are labeled by elements
+$\sigma\in\SS_N$ of the symmetric group $\SS_N$ of order $N$.  We
+refer to simultaneous eigenstates of all $\Pi_\sigma\in\Pi$ as {\it
+  collective states}.  The subset of collective states that have
+eigenvalue 1 with respect to all permutation operators
+$\Pi_\sigma\in\Pi$ is the {\it maximally symmetric collective (MSC)}
+manifold, which is precisely the Dicke manifold in the case of
+two-level subsystems ($n=2$).  The MSC manifold is spanned by a basis
+$\set{\ket{\v m}}$ of MSC states that can be uniquely identified by a
+list of integers $\v m=\p{m_0,m_1,\cdots,m_{n-1}}\in\ZZ_{N+1}^n$,
+where the integer $m_j\in\ZZ_{N+1}$ ($0\le m_j\le N$) indicates the
+total occupation number of subsystem state $j$, and $\sum_j m_j=N$.
+The Dicke manifold of $N$ qubits, for example, is spanned by the Dicke
+states $\set{\ket{m_0,m_1}}$, where $m_0+m_1=N$ and the integer $m_0$
+($m_1$) indicates the occupation number of the qubit state $\ket{0}$
+($\ket{1}$).
+
+The MSC manifold has two important properties that concern us.  First,
+the MSC manifold is closed under the action of unitaries generated by
+linear combinations collective drive operators $\D_{LM}$.  Second, the
+MSC manifold is precisely the ground-state manifold of the collective
+SU($n$)-symmetric interaction Hamiltonian $H_{\t{int}}$ in
+\eqref{eq:H_int_spin} or \eqref{eq:H_int_drive}.  We prove these
+properties of the MSC manifold in Appendix \ref{sec:MSC_proofs}.
 
@@
 \eqref{eq:H_int_spin} takes an identical form in any orthonormal basis
-for $\B_n$, i.e.~the space of operators on the Hilbert space of an
-$n$-level system, equipped with the trace (Hilbert-Schmidt) inner
-product
+for $\B\p{\H_n}$, i.e.~the space of linear operators on the Hilbert
+space $\H_n$ of an $n$-level system, equipped with the trace
+(Hilbert-Schmidt) inner product
 \begin{align}
@@
 To maintain generality, we choose an arbitrary basis
-$X\equiv\set{X_j}$ satisfying the orthonormality condition
+$X\equiv\set{X_j}$ for $\B\p{\H_n}$ that satisfies the orthonormality
+condition
 \begin{align}
@@
 implies that we can resolve the identity (super-)operator $\I$ with
-$\I\oket{\O}=\oket{\O}$ for any $\oket\O\in\B_n$ as
+$\I\oket{\O}=\oket{\O}$ for any $\oket\O\in\B\p{\H_n}$ as
 \begin{align}
@@
 \end{align}
-where we used the fact that $\set{X_j^\dag}$ is also an orthonormal
-basis for $\B_n$, satisfying the same orthonormality condition as
-$\set{X_j}$; these two bases are transformed into each other by
+where we used the fact that $X^\dag\equiv\set{X_j^\dag}$ is also an
+orthonormal basis for $\B\p{\H_n}$, satisfying the same orthonormality
+condition as $X$; these two bases are transformed into each other by
 unitary $\sum_j\oop{X_j^\dag}{X_j}/\N_X$.
@@
 \end{align}
-Choosing a different basis $Y\equiv\set{Y_j}$ for $\B_n$ with
+Choosing a different basis $Y\equiv\set{Y_j}$ for $\B\p{\H_n}$ with
 corresponding norm $\N_Y$, we can expand
@@
 The collective interaction Hamiltonian in \eqref{eq:H_int_spin} thus
-takes an identical form in any orthonormal basis for the space $\B_n$
-of operators on the Hilbert space of an $n$-level system, up to
-appropriate re-scaling factors $\N_X/\N_Y$.
+takes an identical form in any orthonormal basis for the space
+$\B\p{\H_n}$ of linear operators on the Hilbert space $\H_n$ of an
+$n$-level system, up to appropriate re-scaling factors $\N_X/\N_Y$.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Key properties of the maximally-symmetric collective
+  manifold}
+\label{sec:MSC_proofs}
+
+Here we prove two important properties of the maximally-symmetric
+collective (MSC) manifold defined in Section \ref{sec:classifying}:
+\begin{enumerate*}
+\item \label{pt:MSC_closure} that the MSC manifold is closed under the
+  action of unitaries generated by linear combinations collective
+  drive operators $\D_{LM}$; and
+\item \label{pt:MSC_ground} that the MSC manifold is precisely the
+  ground-state manifold of the collective SU($n$)-symmetric
+  interaction Hamiltonian $H_{\t{int}}$ in \eqref{eq:H_int_spin} or
+  \eqref{eq:H_int_drive}.
+\end{enumerate*}
+Our starting ingredients are an $n$-dimensional Hilbert space $\H_n$,
+its $N$-fold tensor product $\H_n^{\otimes N}$, and the set
+$\Pi\equiv\set{\Pi_\sigma}$ of operators that permute the tensor
+factors of $\H_n^{\otimes N}$.  Permutation operators $\Pi_\sigma$ are
+labeled by elements $\sigma\in\SS_N$ of the symmetric group $\SS_N$ of
+order $N$.  The MSC sub-manifold of $\H_n^{\otimes N}$ is defined by
+the set of states that have eigenvalue $1$ with respect to all
+permutation operators $\Pi_\sigma\in\Pi$.
+
+Given an orthonormal basis $X=\set{X_j}$ for the space $\B\p{\H_n}$ of
+linear operators on $\H_n$, we define the collective operator
+$\X_j\in\B\p{\H_n^{\otimes N}}$ by
+\begin{align}
+  \X_j \equiv \sum_{p\in\ZZ_N} X_j^{(p)},
+\end{align}
+where $\ZZ_N\equiv\set{0,1,\cdots,N-1}$ and $X_j^{(p)}$ acts with
+$X_j$ on tensor factor $p$ of $\H_n^{\otimes N}$ and trivially on all
+other tensor factors, i.e.
+\begin{align}
+  X_j^{(p)} \equiv
+  \1^{\otimes p} \otimes X_j \otimes \1^{\otimes\p{N-p-1}}.
+\end{align}
+We denote a vector of all collective operators in $\X\equiv\set{\X_j}$
+by $\v\X$, i.e.~$\v\X\equiv\p{\X_j:\X_j\in\X}$.
+
+Given an arbitrary MSC state
+$\ket{\psi_{\t{MSC}}}\in\H_n^{\otimes N}$, we can use the fact that
+all permutation operators $\Pi_\sigma\in\Pi$ are unitary to expand
+\begin{align}
+  \Pi_\sigma e^{i\v\theta\c\v\X} \ket{\psi_{\t{MSC}}}
+  = \Pi_\sigma e^{i\v\theta\c\v\X}
+  \Pi_\sigma^\dag \Pi_\sigma \ket{\psi_{\t{MSC}}}
+  = e^{i\v\theta\c\Pi_\sigma\v\X\Pi_\sigma^\dag}
+  \Pi_\sigma \ket{\psi_{\t{MSC}}}
+  = e^{i\v\theta\c\Pi_\sigma\v\X\Pi_\sigma^\dag} \ket{\psi_{\t{MSC}}},
+\end{align}
+where
+$\Pi_\sigma\v\X\Pi_\sigma^\dag \equiv \p{\Pi_\sigma\X_j\Pi_\sigma^\dag
+  : \X_j\in\X}$, and
+\begin{align}
+  \Pi_\sigma \X_j \Pi_\sigma^\dag
+  = \sum_{p\in\ZZ_N} \Pi_\sigma X_j^{(p)} \Pi_\sigma^\dag
+  = \sum_{p\in\ZZ_N} X_j^{(\sigma\c p)}
+  = \sum_{p\in\ZZ_N} X_j^{(p)},
+\end{align}
+which implies that $\Pi_\sigma\v\X\Pi_\sigma^\dag=\v\X$, and in turn
+\begin{align}
+  \Pi_\sigma e^{i\v\theta\c\v\X} \ket{\psi_{\t{MSC}}}
+  = e^{i\v\theta\c\v\X} \ket{\psi_{\t{MSC}}}.
+\end{align}
+We thus find that $e^{i\v\theta\c\v\X} \ket{\psi_{\t{MSC}}}$ is itself
+an MSC state, as it has eigenvalue 1 with respect to arbitrary
+permutation operators $\Pi_\sigma\in\Pi$, thereby proving claim
+\ref{pt:MSC_closure} by choice of the orthonormal basis $\set{D_{LM}}$
+for the space $\B\p{\H_n}$ of linear operators on $\H_n$.
+
+[Proof of claim \ref{pt:MSC_ground} to come]
+
 \bibliography{multilevel_spin_notes.bib}

### 2019-10-17 collective_sun/multilevel_spin_notes.tex
capitalized all single-body spin operators, added expression of the interaction Hamiltonian in terms of permutation operators

@@
 \newcommand{\I}{\mathcal{I}}
+\newcommand{\M}{\mathcal{M}}
 \newcommand{\N}{\mathcal{N}}
@@
 
+\newcommand{\MSC}{\text{MSC}}
+
 \DeclareMathOperator{\sign}{sign}
@@
 \setlist{nolistsep} % more compact spacing between environments
+
+%%% text markup
+\usepackage{color} % text color
+\newcommand{\red}[1]{{\color{red} #1}}
 
@@
 divided by the number of lattice sites $L$.  Defining the spin
-operators $s_{\mu\nu}^{(p)}\equiv c_{p\mu}^\dag c_{p\nu}$ for each
+operators $S_{\mu\nu}^{(p)}\equiv c_{p\mu}^\dag c_{p\nu}$ for each
 quasi-momentum $p$, we can alternately write
@@
   H_{\t{int}} = \f{u}{2} \sum_{p,q,\mu,\nu}
-  \p{s_{\mu\mu}^{(p)} s_{\nu\nu}^{(q)} - s_{\mu\nu}^{(p)} s_{\nu\mu}^{(q)}}
-  = \f{u}{2}\p{N^2 - \sum_{p,q,\mu,\nu} s_{\mu\nu}^{(p)} s_{\nu\mu}^{(q)}},
+  \p{S_{\mu\mu}^{(p)} S_{\nu\nu}^{(q)} - S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}}
+  = \f{u}{2}\p{N^2 - \sum_{p,q,\mu,\nu}
+    S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}},
 \end{align}
@@
   H_{\t{int}}
-  = - \f{u}{2} \sum_{p,q,\mu,\nu} s_{\mu\nu}^{(p)} s_{\nu\mu}^{(q)}.
+  = - \f{u}{2} \sum_{p,q,\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}.
 \end{align}
 In the case of SU(2), at this point we would expand the spin operators
-$s_{\mu\nu}^{(p)}$ in terms of Pauli operators in order to convert
+$S_{\mu\nu}^{(p)}$ in terms of Pauli operators in order to convert
 $H_{\t{int}}$ into an SU(2) spin Hamiltonian\cite{he2019engineering}.
 To generalize this procedure to the case of SU($n$), we first define a
-vector $\v s^{(p)}$ of all spin operators $s_{\mu\nu}^{(p)}$, and
+vector $\v S^{(p)}$ of all spin operators $S_{\mu\nu}^{(p)}$, and
 write
@@
   H_{\t{int}}
-  = - \f{u}{2} \sum_{p,q} {\v s^{(p)}}^\dag \v s^{(q)}
+  = - \f{u}{2} \sum_{p,q} {\v S^{(p)}}^\dag \c \v S^{(q)}
   = - \f{u}{2} \v\S^\dag \c \v\S,
   &&
-  \v\S\equiv \sum_p \v s^{(p)}.
+  \v\S\equiv \sum_p \v S^{(p)}.
   \label{eq:H_int_spin}
@@
 $\B\p{\H_n}$ and its dual space $\B\p{\H_n}^*$.  Similarly to the set
-of all spin operators $s_{\mu\nu}\equiv\op{\mu}{\nu}$ on the
+of all spin operators $S_{\mu\nu}\equiv\op{\mu}{\nu}$ on the
 single-particle Hilbert space at each quasi-momentum, all of the bases
@@
   operators}\cite{hioe1981level, bertlmann2008bloch}, defined for
-$j,k,\ell\in\set{0,1,\cdots,n-1}$ with $k<j$ and $\ell>0$ by
+$j,k,\ell\in\set{0,1,\cdots,n-1}$ with $j<k$ and $\ell>0$ by
 \begin{align}
@@
 \begin{align}
-  \lambda_{jk,+} \equiv \op{j}{k} + \op{k}{j},
-  &&
-  \lambda_{jk,-} \equiv i\p{\op{j}{k} - \op{k}{j}},
+  \lambda_{jk,\x} \equiv \op{j}{k} + \op{k}{j},
+  &&
+  \lambda_{jk,\y} \equiv i\p{\op{j}{k} - \op{k}{j}},
   \label{eq:GGM_off_diag}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Classifying collective states}
-\label{sec:classifying}
+\section{Collective states}
+\label{sec:collective}
 
 Given a quantum system composed of $N$ identical (but distinguishable)
-subsystems, e.g.~an array of $n$-level nuclear spins, we identify the
-set $\Pi\equiv\set{\Pi_\sigma}$ of operators that permute subsystems.
-Permutation operators $\Pi_\sigma$ are labeled by elements
-$\sigma\in\SS_N$ of the symmetric group $\SS_N$ of order $N$.  We
-refer to simultaneous eigenstates of all $\Pi_\sigma\in\Pi$ as {\it
-  collective states}.  The subset of collective states that have
-eigenvalue 1 with respect to all permutation operators
-$\Pi_\sigma\in\Pi$ is the {\it maximally symmetric collective (MSC)}
-manifold, which is precisely the Dicke manifold in the case of
-two-level subsystems ($n=2$).  The MSC manifold is spanned by a basis
-$\set{\ket{\v m}}$ of MSC states that can be uniquely identified by a
-list of integers $\v m=\p{m_0,m_1,\cdots,m_{n-1}}\in\ZZ_{N+1}^n$,
-where the integer $m_j\in\ZZ_{N+1}$ ($0\le m_j\le N$) indicates the
-total occupation number of subsystem state $j$, and $\sum_j m_j=N$.
-The Dicke manifold of $N$ qubits, for example, is spanned by the Dicke
-states $\set{\ket{m_0,m_1}}$, where $m_0+m_1=N$ and the integer $m_0$
-($m_1$) indicates the occupation number of the qubit state $\ket{0}$
-($\ket{1}$).
-
-The MSC manifold has two important properties that concern us.  First,
-the MSC manifold is closed under the action of unitaries generated by
-linear combinations collective drive operators $\D_{LM}$.  Second, the
-MSC manifold is precisely the ground-state manifold of the collective
-SU($n$)-symmetric interaction Hamiltonian $H_{\t{int}}$ in
-\eqref{eq:H_int_spin} or \eqref{eq:H_int_drive}.  We prove these
-properties of the MSC manifold in Appendix \ref{sec:MSC_proofs}.
+$n$-level subsystems, e.g.~an array of $n$-level nuclear spins, we
+identify the set $\Pi\equiv\set{\Pi_\sigma}$ of operators that permute
+subsystems, labeled by corresponding elements $\sigma\in\SS_N$ of the
+symmetric group $\SS_N$ of order $N$.  We refer to simultaneous
+eigenstates of all $\Pi_\sigma\in\Pi$ as {\it collective states}.  The
+subset of collective states that have eigenvalue $1$ with respect to
+all permutation operators $\Pi_\sigma\in\Pi$ is the {\it maximally
+  symmetric collective (MSC)} manifold $\M_\MSC$, which is precisely
+the Dicke manifold in the case of two-level subsystems ($n=2$).  The
+MSC manifold is spanned by a basis $\set{\ket{\v m}}$ of MSC states
+that can be uniquely identified by a list of integers
+$\v m=\p{m_0,m_1,\cdots,m_{n-1}}\in\ZZ_{N+1}^n$, where $m_k$ indicates
+the total occupation number of subsystem state $k$, and
+$\sum_k m_k=N$.  The Dicke manifold of $N$ qubits, for example, is
+spanned by the Dicke states $\set{\ket{m_0,m_1}}$, where $m_0+m_1=N$
+and the integer $m_0$ ($m_1$) indicates the occupation number of the
+qubit state $\ket{0}$ ($\ket{1}$).  The dimension of the MSC manifold
+is determined by the number of ways to assign $N$ sub-systems to $n$
+states, which is ${N+n-1 \choose n-1}$.
+
+The MSC manifold has two key properties.  First, the MSC manifold is
+closed under the action of unitaries generated by collective drive
+operators $\D_{LM}$, as well as linear combinations and products
+thereof.  Closure of the MSC manifold under collective dynamics is a
+straightforward consequence of the permutational symmetry obeyed by
+collective operators.  Second, the MSC manifold is the ground-state
+manifold of the collective SU($n$)-symmetric interaction Hamiltonian
+$H_{\t{int}}$ in \eqref{eq:H_int_spin}.  To illuminate the spectral
+properties of $H_{\t{int}}$, we can expand
+\begin{align}
+  H_{\t{int}}
+  = -\f{u}{2} \sum_{p,q,\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}
+  = -u\sum_{\substack{p<q\\\mu,\nu}} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}
+  - \f{u}{2} \sum_{p,\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(p)}
+  = -u\sum_{p<q} \Pi_{pq} - \f12 N n u
+  \simeq -u\sum_{p<q} \Pi_{pq},
+  \label{eq:H_int_perm}
+\end{align}
+where $\Pi_{pq}\in\Pi$ permutes subsystems $p$ and $q$, and $\simeq$
+denotes equality up to an overall constant with no physical
+consequence.  The form of $H_{\t{int}}$ in \eqref{eq:H_int_perm}
+reveals that its eigenstates can be classified by their symmetries
+under the permutation of subsystems.  In particular, the ground-state
+manifold of $H_{\t{int}}$ consists of states that have eigenvalue 1
+with respect to all pair-wise permutation operators $\Pi_{pq}$, which
+is precisely the MSC manifold $\M_\MSC$.  The interaction energy of
+states $\ket{\psi_\MSC}\in\M_\MSC$ is (up to an overall constant)
+simply $-{ N\choose 2 } u = -\p{1/4}N\p{N-1} u$.
 
@@
 \end{align}
-which essentially consists of two-body interactions ($p\ne q$) and
-single-body terms ($p=q$), respectively of the form
-\begin{align}
-  \sum_j X_j \otimes X_j^\dag,
-  &&
-  \sum_j X_j X_j^\dag.
-\end{align}
-Choosing a different basis $Y\equiv\set{Y_j}$ for $\B\p{\H_n}$ with
-corresponding norm $\N_Y$, we can expand
+which essentially consists of single-body terms ($p=q$) and two-body
+interactions ($p\ne q$) of the form
+\begin{align}
+  \sum_j X_j X_j^\dag,
+  &&
+  \sum_j X_j \otimes X_j^\dag.
+\end{align}
+Choosing an arbitrary orthonormal basis $Y\equiv\set{Y_j}$ for
+$\B\p{\H_n}$ with corresponding norm $\N_Y$, we can expand
 \begin{align}
@@
 \end{align}
-By resolution of the identity, we can therefore find
-\begin{align}
+Denoting an arbitrary bilinear operation by $\odot$, by resolution of
+the identity we therefore find
+\begin{align}
+  \sum_j X_j \odot X_j^\dag
+  = \f1{\N_Y^2} \sum_{j,k,\ell} Y_k \obk{Y_k|X_j}
+  \odot Y_\ell^\dag \obk{X_j|Y_\ell}
+  = \f{\N_X}{\N_Y^2} \sum_{k,\ell} Y_k \odot Y_\ell^\dag
+  \obk{Y_k|\I|Y_\ell}
+  = \f{\N_X}{\N_Y} \sum_k Y_k \odot Y_k^\dag,
+\end{align}
+which in particular implies that
+\begin{align}
+  \sum_j X_j X_j^\dag
+  = \f{\N_X}{\N_Y} \sum_k Y_k Y_k^\dag,
+  &&
   \sum_j X_j \otimes X_j^\dag
-  = \f1{\N_Y^2} \sum_{j,k,\ell} Y_k \obk{Y_k|X_j}
-  \otimes Y_\ell^\dag \obk{X_j|Y_\ell}
-  = \f{\N_X}{\N_Y^2} \sum_{k,\ell} Y_k \otimes Y_\ell^\dag
-  \obk{Y_k|\I|Y_\ell}
   = \f{\N_X}{\N_Y} \sum_k Y_k \otimes Y_k^\dag,
 \end{align}
-and similarly
-\begin{align}
-  \sum_j X_j X_j^\dag
-  = \f1{\N_Y^2} \sum_{j,k,\ell} Y_k Y_\ell^\dag
-  \obk{Y_k|X_j} \obk{X_j|Y_\ell}
-  = \f{\N_X}{\N_Y} \sum_k Y_k Y_k^\dag.
+and in turn
+\begin{align}
+  H = \f{\N_X}{\N_Y} \sum_{p,q,j} Y_j^{(p)} {Y_j^{(q)}}^\dag.
 \end{align}
@@
 \eqref{eq:H_lat_SOC}.  In the basis of spin operators
-$s_{\mu\mu}^{(q)} \equiv c_{q\mu}^\dag c_{q\mu}$, this Hamiltonian
+$S_{\mu\mu}^{(q)} \equiv c_{q\mu}^\dag c_{q\mu}$, this Hamiltonian
 takes the form
@@
   H_{\t{lat}}^{(\phi)}
-  = -J \sum_{q,\mu} \cos\p{qa+\mu\phi} s_{\mu\mu}^{(q)}
+  = -J \sum_{q,\mu} \cos\p{qa+\mu\phi} S_{\mu\mu}^{(q)}
   \label{eq:H_lat_spin}
@@
 \begin{align}
-  \tilde d_{\phi+} \equiv \sum_\mu \cos\p{\mu\phi} s_{\mu\mu},
-  &&
-  \tilde d_{\phi-} \equiv \sum_\mu \sin\p{\mu\phi} s_{\mu\mu},
+  \tilde d_{\phi+} \equiv \sum_\mu \cos\p{\mu\phi} S_{\mu\mu},
+  &&
+  \tilde d_{\phi-} \equiv \sum_\mu \sin\p{\mu\phi} S_{\mu\mu},
 \end{align}
@@
 \begin{align}
-  \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
+  \sum_\mu \cos\p{qa+\mu\phi} S_{\mu\mu}
   = \cos\p{qa} \tilde d_{\phi+} - \sin\p{qa} \tilde d_{\phi-},
@@
 \begin{align}
-  \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
+  \sum_\mu \cos\p{qa+\mu\phi} S_{\mu\mu}
   = \cos\p{qa} \f12 \sum_{L,M} \obk{\tilde d_{\phi+}|D_{LM}} D_{LM}
@@
 \begin{align}
-  \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
+  \sum_\mu \cos\p{qa+\mu\phi} S_{\mu\mu}
   = \cos\p{qa} \f12 \sum_{L\,\t{even}}
@@
 \begin{align}
-  \sum_\mu \cos\p{qa+\mu\phi} s_{\mu\mu}
+  \sum_\mu \cos\p{qa+\mu\phi} S_{\mu\mu}
   = \sum_L w_L\p{qa} A_L^{(\phi)} D_{L,0},
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Key properties of the maximally-symmetric collective
-  manifold}
-\label{sec:MSC_proofs}
-
-Here we prove two important properties of the maximally-symmetric
-collective (MSC) manifold defined in Section \ref{sec:classifying}:
-\begin{enumerate*}
-\item \label{pt:MSC_closure} that the MSC manifold is closed under the
-  action of unitaries generated by linear combinations collective
-  drive operators $\D_{LM}$; and
-\item \label{pt:MSC_ground} that the MSC manifold is precisely the
-  ground-state manifold of the collective SU($n$)-symmetric
-  interaction Hamiltonian $H_{\t{int}}$ in \eqref{eq:H_int_spin} or
-  \eqref{eq:H_int_drive}.
-\end{enumerate*}
-Our starting ingredients are an $n$-dimensional Hilbert space $\H_n$,
-its $N$-fold tensor product $\H_n^{\otimes N}$, and the set
-$\Pi\equiv\set{\Pi_\sigma}$ of operators that permute the tensor
-factors of $\H_n^{\otimes N}$.  Permutation operators $\Pi_\sigma$ are
-labeled by elements $\sigma\in\SS_N$ of the symmetric group $\SS_N$ of
-order $N$.  The MSC sub-manifold of $\H_n^{\otimes N}$ is defined by
-the set of states that have eigenvalue $1$ with respect to all
-permutation operators $\Pi_\sigma\in\Pi$.
-
-Given an orthonormal basis $X=\set{X_j}$ for the space $\B\p{\H_n}$ of
-linear operators on $\H_n$, we define the collective operator
-$\X_j\in\B\p{\H_n^{\otimes N}}$ by
-\begin{align}
-  \X_j \equiv \sum_{p\in\ZZ_N} X_j^{(p)},
-\end{align}
-where $\ZZ_N\equiv\set{0,1,\cdots,N-1}$ and $X_j^{(p)}$ acts with
-$X_j$ on tensor factor $p$ of $\H_n^{\otimes N}$ and trivially on all
-other tensor factors, i.e.
-\begin{align}
-  X_j^{(p)} \equiv
-  \1^{\otimes p} \otimes X_j \otimes \1^{\otimes\p{N-p-1}}.
-\end{align}
-We denote a vector of all collective operators in $\X\equiv\set{\X_j}$
-by $\v\X$, i.e.~$\v\X\equiv\p{\X_j:\X_j\in\X}$.
-
-Given an arbitrary MSC state
-$\ket{\psi_{\t{MSC}}}\in\H_n^{\otimes N}$, we can use the fact that
-all permutation operators $\Pi_\sigma\in\Pi$ are unitary to expand
-\begin{align}
-  \Pi_\sigma e^{i\v\theta\c\v\X} \ket{\psi_{\t{MSC}}}
-  = \Pi_\sigma e^{i\v\theta\c\v\X}
-  \Pi_\sigma^\dag \Pi_\sigma \ket{\psi_{\t{MSC}}}
-  = e^{i\v\theta\c\Pi_\sigma\v\X\Pi_\sigma^\dag}
-  \Pi_\sigma \ket{\psi_{\t{MSC}}}
-  = e^{i\v\theta\c\Pi_\sigma\v\X\Pi_\sigma^\dag} \ket{\psi_{\t{MSC}}},
-\end{align}
-where
-$\Pi_\sigma\v\X\Pi_\sigma^\dag \equiv \p{\Pi_\sigma\X_j\Pi_\sigma^\dag
-  : \X_j\in\X}$, and
-\begin{align}
-  \Pi_\sigma \X_j \Pi_\sigma^\dag
-  = \sum_{p\in\ZZ_N} \Pi_\sigma X_j^{(p)} \Pi_\sigma^\dag
-  = \sum_{p\in\ZZ_N} X_j^{(\sigma\c p)}
-  = \sum_{p\in\ZZ_N} X_j^{(p)},
-\end{align}
-which implies that $\Pi_\sigma\v\X\Pi_\sigma^\dag=\v\X$, and in turn
-\begin{align}
-  \Pi_\sigma e^{i\v\theta\c\v\X} \ket{\psi_{\t{MSC}}}
-  = e^{i\v\theta\c\v\X} \ket{\psi_{\t{MSC}}}.
-\end{align}
-We thus find that $e^{i\v\theta\c\v\X} \ket{\psi_{\t{MSC}}}$ is itself
-an MSC state, as it has eigenvalue 1 with respect to arbitrary
-permutation operators $\Pi_\sigma\in\Pi$, thereby proving claim
-\ref{pt:MSC_closure} by choice of the orthonormal basis $\set{D_{LM}}$
-for the space $\B\p{\H_n}$ of linear operators on $\H_n$.
-
-[Proof of claim \ref{pt:MSC_ground} to come]
-
 \bibliography{multilevel_spin_notes.bib}

### 2019-10-17 collective_sun/multilevel_spin_notes.tex
moved the discussion of collective states to a subsection on the interaction Hamiltonian

@@
 
-\newcommand{\MSC}{\text{MSC}}
+\newcommand{\FS}{\text{FS}}
 
@@
 
-\title{Multi-level generalizations of collective spin dynamics and
+\title{Multilevel generalizations of collective spin dynamics and
   squeezing}%
@@
 \end{align}
-In the remainder of this section, we identify various bases for the
-space of linear operators on the Hilbert space $\H_n$ of an $n$-level
+In the bulk of this section, we identify various bases for the space
+of linear operators on the Hilbert space $\H_n$ of an $n$-level
 system.  For brevity, we denote this space of operators by
@@
 phase-space displacement in a discrete space.  As our present work is
-motivated by multi-level systems that are realized with nuclear spin
+motivated by multilevel systems that are realized with nuclear spin
 (angular momentum) degrees of freedom, however, the transition and
@@
 will not discuss Weyl operators any further.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Subsystem permutations and the fully symmetric manifold}
+\label{sec:perm_ops}
+
+For our final re-expression of collective SU($n$)-symmetric
+interactions, we abandon the use of single-spin operators entirely to
+express the interaction Hamiltonian $H_{\t{int}}$ in terms of
+operators that permute tensor factors (subsystems) of a Hilbert space
+with an $N$-fold tensor product structure (e.g.~an array of nuclear
+spins).  Specifically, we denote the Hilbert space of an $n$-level
+system by $\H_n$, denote its $N$-fold tensor product by
+$\H_n^{\otimes N}$, and identify the set $\Pi\equiv\set{\Pi_\sigma}$
+of operators that permute tensor factors of $\H_n^{\otimes N}$
+according to elements $\sigma\in\SS_N$ of the symmetric group of order
+$N$.  We then expand
+\begin{align}
+  H_{\t{int}}
+  = -\f{u}{2} \sum_{p,q,\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}
+  = -u\sum_{\substack{p<q\\\mu,\nu}} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}
+  - \f{u}{2} \sum_{p,\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(p)},
+  \label{eq:H_int_perm_start}
+\end{align}
+where the first sum,
+\begin{align}
+  \sum_{\substack{p<q\\\mu,\nu}} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}
+  = \sum_{\substack{p<q\\\mu,\nu}} \op{\mu}{\nu}^{(p)} \op{\nu}{\mu}^{(q)}
+  = \sum_{\substack{p<q\\\mu,\nu}} \op{\mu\nu}{\nu\mu}^{(p,q)}
+  = \sum_{p<q} \Pi_{pq}
+  \label{eq:spin_perm}
+\end{align}
+is simply a sum over all permutations $\Pi_{pq}$ of subsystems $p$ and
+$q$; and
+\begin{align}
+  \sum_{p,\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(p)}
+  = \sum_{p,\mu,\nu} \op{\mu}{\nu}^{(p)} \op{\nu}{\mu}^{(p)}
+  = n \sum_{p,\mu} \op{\mu}{\mu}^{(p)}
+  = n \sum_p \1^{(p)}
+  = n N
+  \label{eq:spin_const}
+\end{align}
+is merely a constant.  Substituting \eqref{eq:spin_perm} and
+\eqref{eq:spin_const} into \eqref{eq:H_int_perm_start} yields
+\begin{align}
+  H_{\t{int}} = -u \sum_{p<q} \Pi_{pq} - \f12 N n u
+  \simeq -u\sum_{p<q} \Pi_{pq},
+  \label{eq:H_int_perm}
+\end{align}
+where $\simeq$ denotes equality up to an overall constant with no
+physical consequence.  The form of the interaction Hamiltonian
+$H_{\t{int}}$ makes it evident that its ground-state manifold
+$\M_{\FS}$ is the set of {\it fully symmetric} (FS) states that have
+eigenvalue $1$ with respect to all pairwise permutation operators
+$\Pi_{pq}$, and thereby also all $\Pi_\sigma\in\Pi$.  The fully
+symmetric manifold $\M_\FS\subset\H_n^{\otimes N}$ is spanned by a
+basis $\set{\ket{\v m}}$ of fully symmetric states that can be
+uniquely identified by a list of integers
+$\v m\equiv\p{m_0,m_1,\cdots,m_{n-1}}\in\ZZ_{N+1}^n$, where $m_s$
+indicates the total occupation number of subsystem state $s$, and
+$\sum_sm_s=N$.  In the case of $n=2$, the fully symmetric manifold
+$\M_\FS$ is precisely the Dicke manifold\cite{dicke1954coherence} of
+$N$ qubits, spanned by Dicke states $\set{\ket{m_0,m_1}}$ for which
+$m_0+m_1=N$ and the integer $m_0$ ($m_1$) indicates the total
+occupation number of qubit state $\ket{0}$ ($\ket{1}$).  The dimension
+of fully symmetric manifold $\M_\FS$ is determined by the number of
+ways to assign $N$ subsystems to $n$ states, which is
+${N+n-1 \choose n-1}$, and energy of $\ket{\psi_\FS}\in\M_\FS$ with
+respect to the interaction Hamiltonian $H_{\t{int}}$ in
+\eqref{eq:H_int_perm} is $-u{N \choose 2}=-\p{1/4}N\p{N-1}u$.
+
+In addition to the fact that the fully symmetric manifold $\M_\FS$ is
+ground-state manifold of the interaction Hamltonian $H_{\t{int}}$, an
+important feature of $\M_\FS$ is that it is closed under the action of
+unitaries generated by collective operators of the form
+$\Q=\sum_pQ^{(p)}$, as well as linear combinations and products
+thereof.  Closure of $\M_\FS$ under collective dynamics is a
+straightforward consequence of the permutational symmetry obeyed by
+collective operators.
 
@@
 a lattice site, and $\phi$ is the relative phase of the drive between
-adjacent lattice sites.  The multi-level analogue of
+adjacent lattice sites.  The multilevel analogue of
 \eqref{eq:H_drive_2} is
@@
 \end{align}
-Spin-orbit coupling with multi-level systems thus allows for the
+Spin-orbit coupling with multilevel systems thus allows for the
 simulation of synthetic gauge fields with large spin ($L>1$).  As we
@@
 
-Rather than coupling the states of a multi-level system directly with
-a single driving laser, as in Section \ref{sec:drive_SOC}, we now
+Rather than coupling the states of a multilevel system directly with a
+single driving laser, as in Section \ref{sec:drive_SOC}, we now
 consider coupling through an off-resonant Raman transition to a highly
@@
 the drives $\D_{2,1}$ and $\D_{2,2}$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Collective states}
-\label{sec:collective}
-
-Given a quantum system composed of $N$ identical (but distinguishable)
-$n$-level subsystems, e.g.~an array of $n$-level nuclear spins, we
-identify the set $\Pi\equiv\set{\Pi_\sigma}$ of operators that permute
-subsystems, labeled by corresponding elements $\sigma\in\SS_N$ of the
-symmetric group $\SS_N$ of order $N$.  We refer to simultaneous
-eigenstates of all $\Pi_\sigma\in\Pi$ as {\it collective states}.  The
-subset of collective states that have eigenvalue $1$ with respect to
-all permutation operators $\Pi_\sigma\in\Pi$ is the {\it maximally
-  symmetric collective (MSC)} manifold $\M_\MSC$, which is precisely
-the Dicke manifold in the case of two-level subsystems ($n=2$).  The
-MSC manifold is spanned by a basis $\set{\ket{\v m}}$ of MSC states
-that can be uniquely identified by a list of integers
-$\v m=\p{m_0,m_1,\cdots,m_{n-1}}\in\ZZ_{N+1}^n$, where $m_k$ indicates
-the total occupation number of subsystem state $k$, and
-$\sum_k m_k=N$.  The Dicke manifold of $N$ qubits, for example, is
-spanned by the Dicke states $\set{\ket{m_0,m_1}}$, where $m_0+m_1=N$
-and the integer $m_0$ ($m_1$) indicates the occupation number of the
-qubit state $\ket{0}$ ($\ket{1}$).  The dimension of the MSC manifold
-is determined by the number of ways to assign $N$ sub-systems to $n$
-states, which is ${N+n-1 \choose n-1}$.
-
-The MSC manifold has two key properties.  First, the MSC manifold is
-closed under the action of unitaries generated by collective drive
-operators $\D_{LM}$, as well as linear combinations and products
-thereof.  Closure of the MSC manifold under collective dynamics is a
-straightforward consequence of the permutational symmetry obeyed by
-collective operators.  Second, the MSC manifold is the ground-state
-manifold of the collective SU($n$)-symmetric interaction Hamiltonian
-$H_{\t{int}}$ in \eqref{eq:H_int_spin}.  To illuminate the spectral
-properties of $H_{\t{int}}$, we can expand
-\begin{align}
-  H_{\t{int}}
-  = -\f{u}{2} \sum_{p,q,\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}
-  = -u\sum_{\substack{p<q\\\mu,\nu}} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}
-  - \f{u}{2} \sum_{p,\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(p)}
-  = -u\sum_{p<q} \Pi_{pq} - \f12 N n u
-  \simeq -u\sum_{p<q} \Pi_{pq},
-  \label{eq:H_int_perm}
-\end{align}
-where $\Pi_{pq}\in\Pi$ permutes subsystems $p$ and $q$, and $\simeq$
-denotes equality up to an overall constant with no physical
-consequence.  The form of $H_{\t{int}}$ in \eqref{eq:H_int_perm}
-reveals that its eigenstates can be classified by their symmetries
-under the permutation of subsystems.  In particular, the ground-state
-manifold of $H_{\t{int}}$ consists of states that have eigenvalue 1
-with respect to all pair-wise permutation operators $\Pi_{pq}$, which
-is precisely the MSC manifold $\M_\MSC$.  The interaction energy of
-states $\ket{\psi_\MSC}\in\M_\MSC$ is (up to an overall constant)
-simply $-{ N\choose 2 } u = -\p{1/4}N\p{N-1} u$.
 

### 2019-10-19 collective_sun/multilevel_spin_notes.tex
fixed some sign errors in the appendix on drive operators

@@
 \begin{align}
-  D_{LM} \equiv \f{\tau_M}{\sqrt{2}}
-  \sp{T_{L\abs{M}} + \sign\p{M}T_{L\abs{M}}^\dag},
-  &&
-  \tau_M \equiv
+  D_{LM} \equiv \f{\eta_M}{\sqrt{2}}
+  \sp{T_{L\abs{M}} + \sign\p{M} T_{L\abs{M}}^\dag},
+  &&
+  \eta_M \equiv
   \begin{cases}
@@
 \end{align}
-The SU($n$) symmetry of collective interactions implies that the
-interaction Hamiltonian $H_{\t{int}}$ in \eqref{eq:H_int_drive} is
-invariant under this gauge transformation, but the single-particle
-Hamiltonian in \eqref{eq:H_lat} transforms as (see Appendix
-\ref{sec:lat_drive})
+The SU($n$) symmetry of collective interactions considered in this
+work implies that the interaction Hamiltonian $H_{\t{int}}$ in
+\eqref{eq:H_int_perm} is unaffected by this gauge transformation.  The
+single-particle Hamiltonian $H_{\t{lat}}$ in \eqref{eq:H_lat},
+however, transforms as (see Appendix \ref{sec:lat_drive})
 \begin{align}
@@
 \end{align}
-Spin-orbit coupling with multilevel systems thus allows for the
+The multilevel spin-orbit coupling (SOC) exhibited by the
+singe-particle Hamiltonian $H_{\t{lat}}^{(\phi)}$ thus allows for the
 simulation of synthetic gauge fields with large spin ($L>1$).  As we
-will see in the following section, multi-laser drives can also
+will see in Section \ref{sec:drive_raman}, multi-laser drives can also
 simulate synthetic large-spin fields with non-zero spin projection
 ($M\ne0$) onto a quantization axis.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Perturbative treatment of spin-orbit coupling}
+\label{sec:pert_SOC}
+
+When SOC is weak,
+
+\vspace{3cm}
 
@@
 \end{align}
+
+\vspace{3cm}
 
@@
 the drives $\D_{2,1}$ and $\D_{2,2}$.
-
-\vspace{3cm}
 
@@
 \end{align}
+or, equivalently,
+\begin{align}
+  D_{LM} \equiv \f{\tau_M}{\sqrt{2}}
+  \sp{T_{LM} + \sign\p{M} T_{LM}^\dag},
+  &&
+  \tau_M \equiv
+  \begin{cases}
+    \sqrt{2} & M = 0 \\
+    \p{-1}^M \sign\p{M}^{M+1/2} & M \ne 0
+  \end{cases},
+\end{align}
 which implies that
@@
     \tau_M^* & M \ne 0
-  \end{cases}.
+  \end{cases},
 \end{align}
@@
 \end{align}
-with operators
+where in terms of
+\begin{align}
+  \epsilon_M \equiv \p{-1}^M \sign\p{M},
+\end{align}
+we have
 \begin{align}

### 2019-10-23 collective_sun/multilevel_spin_notes.tex
added derivation of perturbative effect of SOC on the fully symmetric manifold

@@
 \newcommand{\D}{\mathcal{D}}
+\newcommand{\E}{\mathcal{E}}
 \renewcommand{\H}{\mathcal{H}}
@@
 \renewcommand{\O}{\mathcal{O}}
+\renewcommand{\P}{\mathcal{P}}
 \newcommand{\Q}{\mathcal{Q}}
+\newcommand{\R}{\mathcal{R}}
 \newcommand{\T}{\mathcal{T}}
@@
 \newcommand{\X}{\mathcal{X}}
-
+\newcommand{\Z}{\mathcal{Z}}
+
+\newcommand{\EE}{\mathbb{E}}
 \renewcommand{\SS}{\mathbb{S}}
@@
 \DeclareMathOperator{\sign}{sign}
+\let\var\relax
+\DeclareMathOperator{\var}{var}
 
@@
 \begin{align}
-  H_{\t{lat}} = -J \sum_{q,\mu} \cos\p{qa} c_{q\mu}^\dag c_{q\mu},
+  H_{\t{lat}} = -t \sum_{q,\mu} \cos\p{qa} c_{q\mu}^\dag c_{q\mu},
   \label{eq:H_lat}
 \end{align}
-where $J$ is the nearest-neighbor tunneling rate, $a$ is the spacing
+where $t$ is the nearest-neighbor tunneling rate, $a$ is the spacing
 between neighboring lattice sites, and $c_{q\mu}$ is the fermionic
@@
 approximation, collective two-body SU($n$)-symmetric interactions
-between these fermions take form\cite{perlin2019effective}
+between these fermions take form
 \begin{align}
@@
 $\Pi_{pq}$, and thereby also all $\Pi_\sigma\in\Pi$.  The fully
-symmetric manifold $\M_\FS\subset\H_n^{\otimes N}$ is spanned by a
-basis $\set{\ket{\v m}}$ of fully symmetric states that can be
-uniquely identified by a list of integers
-$\v m\equiv\p{m_0,m_1,\cdots,m_{n-1}}\in\ZZ_{N+1}^n$, where $m_s$
-indicates the total occupation number of subsystem state $s$, and
-$\sum_sm_s=N$.  In the case of $n=2$, the fully symmetric manifold
+symmetric manifold $\M_\FS$ is spanned by a basis $\set{\ket{\v m}}$
+of fully symmetric states that can be uniquely identified by a list of
+integers $\v m\equiv\p{m_0,m_1,\cdots,m_{n-1}}\in\ZZ_{N+1}^n$, where
+$m_s$ indicates the total occupation number of subsystem state $s$,
+and $\sum_sm_s=N$.  In the case of $n=2$, the fully symmetric manifold
 $\M_\FS$ is precisely the Dicke manifold\cite{dicke1954coherence} of
@@
 ways to assign $N$ subsystems to $n$ states, which is
-${N+n-1 \choose n-1}$, and energy of $\ket{\psi_\FS}\in\M_\FS$ with
-respect to the interaction Hamiltonian $H_{\t{int}}$ in
-\eqref{eq:H_int_perm} is $-u{N \choose 2}=-\p{1/4}N\p{N-1}u$.
+${N+n-1 \choose n-1}$, and the energy of a state
+$\ket{\psi_\FS}\in\M_\FS$ with respect to the interaction Hamiltonian
+$H_{\t{int}}$ in \eqref{eq:H_int_perm} is
+$-u{N \choose 2}=-\p{1/4}N\p{N-1}u$.
 
@@
   \to H_{\t{lat}}^{(\phi)}
-  = -J \sum_{q,\mu} \cos\p{qa+\mu\phi} c_{q\mu}^\dag c_{q\mu}
+  = -t \sum_{q,\mu} \cos\p{qa+\mu\phi} c_{q\mu}^\dag c_{q\mu}
   = \sum_{q,L} B_L^{(\phi,q)} D_{L,0}^{(q)},
-  \label{eq:H_lat_SOC}
+  \label{eq:H_lat_SOC_B}
 \end{align}
@@
 \begin{align}
-  B_L^{(\phi,q)} = -J w_L\p{qa} A_L^{(\phi)},
+  B_L^{(\phi,q)} = -t w_L\p{qa} A_L^{(\phi)},
 \end{align}
@@
   \sum_\mu \bk{I\mu;L,0|I\mu} w_L\p{\mu\phi}.
+  \label{eq:A_L_phi}
 \end{align}
@@
 
-When SOC is weak,
-
-\vspace{3cm}
-
-In the regime of weak SOC with $\abs{\phi}\ll1$, by considering
+The single-particle Hamiltonian $H_{\t{lat}}^{(\phi)}$ breaks
+permutational symmetry, thereby coupling states within the fully
+symmetric manifold $\M_\FS$ to states outsides it.  If SOC is
+sufficiently weak, however, transitions outside the fully symmetric
+manifold $\M_\FS$ are energetically suppressed by interactions, and we
+can account for the effect of SOC perturbatively.  In the regime of
+weak SOC with $\abs{\phi}\ll1$, by considering
 \begin{enumerate*}
@@
   &&
-  n_L \equiv \prod_{\ell=-L}^L\p{n-\ell}.
+  n_L \equiv \prod_{\ell=-L}^L\p{n+\ell}.
 \end{align}
@@
 spin-orbit coupled Hamiltonian $H_{\t{lat}}^{(\phi)}$ in
-\eqref{eq:H_lat_SOC}.  In the basis of spin operators
+\eqref{eq:H_lat_SOC_B}.  In the basis of spin operators
 $S_{\mu\mu}^{(q)} \equiv c_{q\mu}^\dag c_{q\mu}$, this Hamiltonian
@@
   H_{\t{lat}}^{(\phi)}
-  = -J \sum_{q,\mu} \cos\p{qa+\mu\phi} S_{\mu\mu}^{(q)}
+  = -t \sum_{q,\mu} \cos\p{qa+\mu\phi} S_{\mu\mu}^{(q)}
   \label{eq:H_lat_spin}
 \end{align}
-Defining the diagonal ``spin-wave'' operators
+Defining the diagonal operators
 \begin{align}
@@
   &&
-  B_L^{(\phi,q)} \equiv -J w_L\p{qa} A_L^{(\phi)}.
+  B_L^{(\phi,q)} \equiv -J
+  w_L\p{qa} A_L^{(\phi)}.
 \end{align}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Weak SOC in the fully symmetric manifold}
+
+Here we derive the effective Hamiltonian induced on the fully
+symmetric manifold $\M_\FS$ by the single-particle SOC Hamiltonian
+$H_{\t{lat}}^{(\phi)}$ in \eqref{eq:H_lat_SOC_B} in the weak SOC limit
+$\abs{\phi}\ll1$.  For brevity, we work in units with the lattice
+spacing $a=1$ throughout this section.  Perturbations $V$ to the
+interaction Hamiltonian $H_{\t{int}}$ induce the following effective
+Hamiltonians on the fully symmetric manifold
+$\M_\FS$\cite{bravyi2011schrieffer, perlin2019effective}:
+\begin{align}
+  H_{\t{eff}}^{(1)} = \P_0 V \P_0,
+  &&
+  H_{\t{eff}}^{(2)} = -\P_0 V \E V \P_0,
+  &&
+  \E \equiv \sum_{E_\varepsilon>0} \f{\op{E_\varepsilon}}{E_\varepsilon},
+  \label{eq:pert}
+\end{align}
+where $\P_0$ is a projector onto the fully symmetric manifold $\M_\FS$
+and $\ket{E_\varepsilon}$ is an eigenstate of the interaction
+Hamiltonian $H_{\t{int}}$ with energy $E_\varepsilon$ relative to the
+interaction energy of fully symmetric states
+$\ket{\psi_\FS}\in\M_\FS$.  The sum over all $E_\varepsilon>0$ is a
+sum over a complete basis of states $\set{\ket{E_\varepsilon}}$ for
+the orthogonal complement of $\M_\FS$.
+
+\subsection{Effective SOC at first order}
+
+The single-particle SOC Hamiltonian is
+\begin{align}
+  H_{\t{lat}}^{(\phi)}
+  = -t \sum_{q,L} w_L\p{q} A_L^{(\phi)} D_{L,0}^{(q)},
+  &&
+  w_L\p{q} \equiv
+  \begin{cases}
+    \cos q & L~\t{even} \\
+    \sin q & L~\t{odd}
+  \end{cases},
+  \label{eq:H_lat_SOC_A}
+\end{align}
+where $A_L^{(\phi)}$ is defined in \eqref{eq:A_L_phi}, and scales as
+$A_L^{(\phi)} \sim \phi^L + O\p{\phi^{L+2}}$ when $\abs{\phi}\ll1$.
+When the SOC angle $\phi$ is sufficiently small in magnitude, the
+effective Hamiltonian induced on the fully symmetric manifold $\M_\FS$
+at first order in the perturbation $H_{\t{lat}}^{(\phi)}$ is
+\begin{align}
+  H_{\t{eff}}^{(1)}
+  = \P_0 H_{\t{lat}}^{(\phi)} \P_0
+  = -t \sum_{q,L} w_L\p{q} A_L^{(\phi)} \P_0 D_{L,0}^{(q)} \P_0,
+\end{align}
+where we can freely change the index of each single-mode drive
+operator $D_{L,0}^{(q)}$ within the fully symmetric manifold to
+re-write
+\begin{align}
+  \P_0 D_{L,0}^{(q)} \P_0 = \f1N \sum_p \P_0 D_{L,0}^{(p)} \P_0
+  = \f1N \P_0 \D_{L,0} \P_0,
+\end{align}
+which implies that
+\begin{align}
+  H_{\t{eff}}^{(1)}
+  = -t \sum_L \EE_q\sp{w_L\p{q}} A_L^{(\phi)} \D_{L,0},
+  &&
+  \EE_q\sp{X\p{q}} \equiv \f1N \sum_q X\p{q}.
+  \label{eq:H_eff_1}
+\end{align}
+Here $\EE_q\sp{X\p{q}}$ is an average of $X\p{q}$ over all occupied
+quasi-momenta $q$.  Note that we have dropped the explicit projection
+operators $\P_0$ in \eqref{eq:H_eff_1} with the understanding that we
+are implicitly restricting all effective dynamics to the fully
+symmetric manifold $\M_\FS$.
+
+\subsection{Effective SOC at second order}
+
+In order to expand the second-order effective Hamiltonian
+$H_{\t{eff}}^{(2)}$ induced on the fully symmetric manifold $\M_\FS$
+by the SOC Hamiltonian $H_{\t{lat}}^{(\phi)}$ when $\abs{\phi}\ll1$,
+we expand $w_L\p{q}$ in \eqref{eq:H_lat_SOC_A} as
+\begin{align}
+  w_L\p{q} = \f12 \sum_{s\in\set{\pm1}} \chi_{Ls} e^{isq},
+  &&
+  \chi_{Ls} \equiv \p{-i}^{L\,\t{mod}\,2} s^L,
+\end{align}
+which implies that
+\begin{align}
+  H_{\t{lat}}^{(\phi)}
+  = -\f{t}{2} \sum_{\substack{L\\s\in\set{\pm1}}}
+  \chi_{Ls} A_L^{(\phi)} \Z_{Ls},
+  &&
+  \Z_{Ks} \equiv \sum_q e^{isq} D_{L,0}^{(q)},
+\end{align}
+and in turn
+\begin{align}
+  H_{\t{eff}}^{(2)}
+  = -\f{t^2}{4} \sum_{\substack{J,K\\r,s\in\set{\pm1}}}
+  \chi_{Jr} \chi_{Ks} A_J^{(\phi)} A_K^{(\phi)}
+  \P_0 \Z_{Jr} \E \Z_{Ks} \P_0,
+  \label{eq:H_eff_2_start}
+\end{align}
+where $\E$ is a sum over excited-state projectors with energetic
+suppression factors, defined in \eqref{eq:pert}.
+
+\subsubsection{Eliminating excited-state projectors in $\E$}
+
+In order to simplify \eqref{eq:H_eff_2_start}, we first note that the
+image of any fully symmetric state $\ket{\psi_{\FS}}\in\M_\FS$ under
+the action of the ``spin-wave'' operator $\Z_{Ks}$ is an eigenvector
+of the interaction Hamiltonian $H_{\t{int}}=-u\sum_{p<q}\Pi_{pq}$.  To
+substantiate this claim, we expand
+\begin{align}
+  \sum_{p<q} \Pi_{pq} \Z_{Ks} \ket{\psi_\FS}
+  = \sum_{\substack{p<q\\k\notin\set{p,q}}} e^{isk}
+  \Pi_{pq} D_{K,0}^{(k)} \ket{\psi_\FS}
+  + \sum_{\substack{p<q\\k\in\set{p,q}}} e^{isk}
+  \Pi_{pq} D_{K,0}^{(k)} \ket{\psi_\FS}.
+\end{align}
+If $k\notin\set{p,q}$, then the permutation operator $\Pi_{pq}$
+commutes with $D_{K,0}^{(k)}$, and in turn acts trivially on the
+permutationally-symmetric state $\ket{\psi_\FS}$, giving us
+\begin{align}
+  \sum_{\substack{p<q\\k\notin\set{p,q}}} e^{isk}
+  \Pi_{pq} D_{K,0}^{(k)} \ket{\psi_\FS}
+  = \sum_k \sum_{\substack{p<q\\p,q\ne k}} e^{isk}
+  D_{K,0}^{(k)} \ket{\psi_\FS}
+  = { N-1 \choose 2 } \Z_{Ks} \ket{\psi_\FS}.
+\end{align}
+If $k\in\set{p,q}$, then we can resolve the identity operator
+$\1=\Pi_{pq}^\dag \Pi_{pq}$ to expand
+\begin{align}
+  \sum_{\substack{p<q\\k\in\set{p,q}}} e^{isk}
+  \Pi_{pq} D_{K,0}^{(k)} \ket{\psi_\FS}
+  = \f12 \sum_{\substack{p\ne q\\k\in\set{p,q}}} e^{isk}
+  \Pi_{pq} D_{K,0}^{(k)} \Pi_{pq}^\dag \Pi_{pq} \ket{\psi_\FS}
+  = \sum_{q\ne k} e^{isk} D_{K,0}^{(q)} \ket{\psi_\FS},
+\end{align}
+where for $s\in\set{\pm1}$ we can simplify the sum
+$\sum_{k\ne q} e^{isk} = -e^{isq}$,\footnote{We assume for now that
+  $\sum_k e^{\pm ik}=0$, which is only valid when all quasi-momenta
+  are occupied, such that the sum $\sum_k$ runs over all angles
+  $k\in\ZZ_N\times2\pi/N$.  We will consider the slightly more
+  complicated case with $\sum_k e^{\pm ik}\ne0$ later in this
+  section.}  so
+\begin{align}
+  \sum_{\substack{p<q\\k\in\set{p,q}}} e^{isk}
+  \Pi_{pq} D_{K,0}^{(k)} \ket{\psi_\FS}
+  = - \sum_q e^{isq} D_{K,0}^{(q)} \ket{\psi_\FS}
+  = - \Z_{Ks} \ket{\psi_\FS}.
+\end{align}
+In total, we have that
+\begin{align}
+  \sum_{p<q} \Pi_{pq} \Z_{Ks} \ket{\psi_\FS}
+  = \sp{{ N-1 \choose 2 } - 1} \Z_{Ks} \ket{\psi_\FS},
+\end{align}
+which implies that $\Z_{Ks}\ket{\psi_\FS}$ is proportional to an
+eigenstate of $H_{\t{int}}$ with interaction energy
+$-u\sp{{ N-1 \choose 2 } - 1}$, which is $Nu$ above the interaction
+energy of $-u{N\choose2}$ of states in the fully symmetric manifold,
+so
+\begin{align}
+  \E \Z_{Ks} \P_0
+  =  \f1{Nu} \Z_{Ks} \P_0
+  = \f1{fU} \Z_{Ks} \P_0,
+  \label{eq:proj_elim}
+\end{align}
+where $\E$ is defined in \eqref{eq:pert} and $f\equiv N/L$ is the
+filling fraction of quasi-momentum modes.  We have thus far simplified
+the effective Hamiltonian in \eqref{eq:H_eff_2_start} to
+\begin{align}
+  H_{\t{eff}}^{(2)}
+  = -\f{t^2}{4fU} \sum_{\substack{J,K\\r,s\in\set{\pm1}}}
+  \chi_{Jr} \chi_{Ks} A_J^{(\phi)} A_K^{(\phi)}
+  \P_0 \Z_{Jr} \Z_{Ks} \P_0,
+  \label{eq:H_eff_2_mid}
+\end{align}
+leaving us the task of evaluating the product
+$\P_0 \Z_{Jr} \Z_{Ks} \P_0$, as well as the sum over $r,s$.
+
+\subsubsection{The case of $r=s$}
+
+To evaluate the product $\P_0 \Z_{Jr} \Z_{Ks} \P_0$ of spin-wave
+operators projected onto the fully symmetric manifold $\M_\FS$, we
+first consider the case of $r=s$, and use the permutational symmetry
+enforced by the projectors $\P_0$ to expand
+\begin{align}
+  \P_0 \Z_{J\pm} \Z_{K\pm} \P_0
+  &= \sum_{p,q} e^{\pm i\p{p+q}} \P_0 D_{J,0}^{(p)} D_{K,0}^{(q)} \P_0 \\
+  &= \sum_p e^{\pm i2p} \P_0 D_{J,0}^{(0)} D_{K,0}^{(0)} \P_0
+  + \sum_{p\ne q} e^{\pm i\p{p+q}} \P_0 D_{J,0}^{(0)} D_{K,0}^{(1)} \P_0,
+\end{align}
+where $\sum_p e^{\pm i2p}=0$, and
+\begin{align}
+  \sum_{p\ne q} e^{\pm i\p{p+q}}
+  = \sum_p e^{\pm ip} \sum_{q\ne p} e^{\pm iq}
+  = \sum_p e^{\pm ip} \p{-e^{\pm ip}}
+  = -\sum_p e^{\pm i2p}
+  = 0,
+\end{align}
+which implies that
+\begin{align}
+  \P_0 \Z_{J\pm} \Z_{K\pm} \P_0 = 0.
+  \label{eq:ZZ_eq}
+\end{align}
+
+\subsubsection{The case of $r\ne s$}
+\label{sec:H_eff_2_end}
+
+For the case of $r\ne s$, we similarly expand
+\begin{align}
+  \P_0 \Z_{J\pm} \Z_{K\mp} \P_0
+  &= \sum_{p,q} e^{\pm i\p{p-q}} \P_0 D_{J,0}^{(p)} D_{K,0}^{(q)} \P_0 \\
+  &= N \P_0 D_{J,0}^{(0)} D_{K,0}^{(0)} \P_0
+  + \sum_{p\ne q} e^{\pm i\p{p-q}} \P_0 D_{J,0}^{(0)} D_{K,0}^{(1)} \P_0,
+\end{align}
+where
+\begin{align}
+  \sum_{p\ne q} e^{\pm i\p{p-q}}
+  = \sum_p e^{\pm ip} \sum_{q\ne p} e^{\mp iq}
+  = \sum_p e^{\pm ip} \p{-e^{\mp ip}}
+  = - \sum_p e^{\pm i\p{p-p}}
+  = - N,
+\end{align}
+so
+\begin{align}
+  \P_0 \Z_{J\pm} \Z_{K\mp} \P_0
+  = N \P_0 D_{J,0}^{(0)} D_{K,0}^{(0)} \P_0
+  - N \P_0 D_{J,0}^{(0)} D_{K,0}^{(1)} \P_0.
+  \label{eq:ZZ_neq_mid}
+\end{align}
+In order to express this product in terms of collective operators, we
+expand
+\begin{align}
+  \P_0 \D_{J,0} \D_{K,0} \P_0
+  = N \P_0 D_{J,0}^{(0)} D_{K,0}^{(0)} \P_0
+  + N \p{N-1} \P_0 D_{J,0}^{(0)} D_{K,0}^{(1)} \P_0,
+\end{align}
+then solve for the last term and substitute the result into
+\eqref{eq:ZZ_neq_mid}, giving us
+\begin{align}
+  \P_0 \Z_{J\pm} \Z_{K\mp} \P_0
+  = \f{N^2}{N-1} \P_0 D_{J,0}^{(0)} D_{K,0}^{(0)} \P_0
+  - \f1{N-1} \P_0 \D_{J,0} \D_{K,0} \P_0.
+\end{align}
+Using the structure constants $f_{\ell_1m_1;\ell_2m_2}^{LM}$ of the
+transition operator algebra (see Appendix \ref{sec:trans_prod}), we
+then expand
+\begin{align}
+  N \P_0 D_{J,0}^{(0)} D_{K,0}^{(0)} \P_0
+  = N \sum_L f_{J,0;K,0}^{L,0} \P_0 D_{L,0}^{(0)} \P_0
+  = \sum_L f_{J,0;K,0}^{L,0} \P_0 \D_{L,0} \P_0,
+\end{align}
+which implies that
+\begin{align}
+  \P_0 \Z_{J\pm} \Z_{K\mp} \P_0
+  = \f{N}{N-1} \sum_L f_{J,0;K,0}^{L,0} \P_0 \D_{L,0} \P_0
+  - \f1{N-1} \P_0 \D_{J,0} \D_{K,0} \P_0.
+  \label{eq:ZZ_neq}
+\end{align}
+
+\subsubsection{Putting it all together}
+
+Having evaluated the product $\P_0 \Z_{Jr} \Z_{Ks} \P_0$ of spin-wave
+operators in \eqref{eq:ZZ_eq} and \eqref{eq:ZZ_neq}, the last
+ingredient needed to simplify the effective Hamiltonian
+$H_{\t{eff}}^{(2)}$ in \eqref{eq:H_eff_2_mid} is an evaluation of the
+sum over $r,s\in\set{\pm1}$.  The product of spin-wave operators
+enforces $r\ne s$, but is otherwise independent of $r,s$, so we need
+only evaluate
+\begin{align}
+  \sum_{s\in\set{\pm1}} \chi_{J,-s} \chi_{Ks}
+  =
+  \begin{cases}
+    2 & J = K \, \t{mod}\, 2 \\
+    0 & \t{otherwise}
+  \end{cases}.
+\end{align}
+Putting everything together, we thus find
+\begin{align}
+  H_{\t{eff}}^{(2)}
+  = \f{t^2}{2\p{N-1}fU}
+  \sum_{J=K\,\t{mod}\,2} A_J^{(\phi)} A_K^{(\phi)}
+  \sp{\D_{J,0} \D_{K,0} - N\sum_L f_{J,0;K,0}^{L,0} \D_{L,0}},
+  \label{eq:H_eff_2_unit}
+\end{align}
+where, as for $H_{\t{eff}}^{(1)}$ in \eqref{eq:H_eff_1}, we drop the
+explicit projectors $\P_0$ in \eqref{eq:H_eff_2_unit} with the
+understanding that we are implicitly restricting all effective
+dynamics to the fully symmetric manifold $\M_\FS$.
+
+Thus far, we have assumed that the sum over quasi-momenta
+$\sum_ke^{\pm ik}=0$, which is only valid when all quasi-momenta are
+occupied, such that the sum $\sum_k$ runs over all angles
+$k\in\ZZ_N\times2\pi/N$.  If not all quasi-momenta are occupied, then
+we can repeat all derivations above with mean-subtracted spin-wave
+operators
+\begin{align}
+  \Z_{Ks} \equiv \sum_q e^{isq} D_{K,0}^{(q)}
+  \to
+  \tilde\Z_{Ks}
+  \equiv \sum_q \p{e^{isq} - \EE_k\sp{e^{isk}}} D_{K,0}^{(q)},
+\end{align}
+where $\EE_k\sp{e^{isk}}$ is an average of $e^{isk}$ over all occupied
+quasi-momenta $k$.  Note that the difference
+$\tilde\Z_{Ks}-\Z_{Ks}=-\EE_k\sp{e^{isk}}\D_{K,0}$ is permutationally
+symmetric, which implies that it only couples fully symmetric states
+$\ket{\psi_\FS}\in\M_\FS$ to fully symmetric states.  In particular,
+due to the fact that $\E$ is a linear combination of projectors onto
+interaction-energy eigenstates strictly outside of $\M_\FS$,
+\begin{align}
+  \P_0 \tilde\Z_{Jr} \E \tilde\Z_{Ks} \P_0
+  = \P_0 \Z_{Jr} \E \Z_{Ks} \P_0,
+\end{align}
+which implies that we can freely take
+$\Z_{Jr},\Z_{Ks}\to\tilde\Z_{Jr},\tilde\Z_{Ks}$ with no consequence
+for the effective Hamiltonian $H_{\t{eff}}^{(2)}$ in
+\eqref{eq:H_eff_2_start}.
+
+Repeating the previous calculations in this section with the
+mean-subtracted spin-wave operators, we still find that, just as in
+\eqref{eq:proj_elim},
+\begin{align}
+  \E \tilde\Z_{Ks} \P_0 = \f1{fU} \tilde\Z_{Ks} \P_0,
+\end{align}
+whereas, in contrast to \eqref{eq:ZZ_eq} and \eqref{eq:ZZ_neq},
+\begin{align}
+  \P_0 \Z_{J\pm} \Z_{K\pm} \P_0
+  = -\f{\var_q\sp{e^{\pm iq}}}{N-1}
+  \sp{\P_0 \D_{J,0} \D_{K,0} \P_0
+    - N \sum_L f_{J,0;K,0}^{L,0} \P_0 \D_{L,0} \P_0},
+\end{align}
+where
+$\var_q\sp{X_q} \equiv \EE_q\sp{\p{X_q-\EE_p\sp{X_p}}^2} =
+\EE_q\sp{X_q^2}-\EE_q\sp{X_q}^2$ is the variance of $X_q$ over all
+occupied quasi-momenta $q$, and
+\begin{align}
+  \P_0 \Z_{J\pm} \Z_{K\pm} \P_0
+  = -\f{1-\EE_p\sp{e^{ip}}\EE_q\sp{e^{-iq}}}{N-1}
+  \sp{\P_0 \D_{J,0} \D_{K,0} \P_0
+    - N \sum_L f_{J,0;K,0}^{L,0} \P_0 \D_{L,0} \P_0}.
+\end{align}
+
+\vspace{3cm}
+
+\red{To do: Work out SU($2$) explicitly as a special case in these
+  notes.  Also, the current second-order derivation assumes unit
+  filling, and needs to be adjusted for the possibility of $f<1$.  I
+  think that in the case of $f<1$ the factor of $1/2$ in
+  \eqref{eq:H_eff_2} turns into $\bk{w_J\p{q}w_K\p{q}}_q$, e.g.~when
+  $J=K=1$, this factor becomes the variance of $\sin\p{q}$ over all
+  occupied modes, as in the known case of SU($2$).  I should also work
+  out explicitly the contributions through second order in $\phi$,
+  which will simplify things somewhat as we will only need $J,K\le2$
+  in $H_{\t{lat}}^{(1)}$, and $J=K=1$ in $H_{\t{lat}}^{(2)}$.  I can
+  also do a third-order calculation of $H_{\t{eff}}^{(3)}$ to extract
+  dependence on $\phi^3$, but that's probably unnecessary / not
+  particularly interesting.}
+
+\vspace{3cm}
+
 \bibliography{multilevel_spin_notes.bib}

### 2019-10-23 collective_sun/multilevel_spin_notes.tex
cleaned up the derivation of an effective weak-SOC hamiltonian

--- 2019-10-19 collective_sun/multilevel_spin_notes.tex
@@
 \DeclareMathOperator{\sign}{sign}
+\DeclareMathOperator{\cov}{cov}
+\let\var\relax
+\DeclareMathOperator{\var}{var}
 
@@
 ground-state manifold of the interaction Hamltonian $H_{\t{int}}$, an
-important feature of $\M_\FS$ is that it is closed under the action of
+important feature of $\M_\FS$ is its closure under the action of
 unitaries generated by collective operators of the form
@@
   A_L^{(\phi)} = c_L \sqrt{n_L}\, \phi^L + O\p{\phi^{L+2}},
+  &&
+  c_L \equiv \p{-1}^{L+\lfloor L/2\rfloor}
+  \f{\sqrt{2L+1}\,L!}{\p{2L+1}!},
+  &&
+  n_L \equiv \prod_{\ell=-L}^L\p{n+\ell}.
   \label{eq:A_L_phi_small}
-\end{align}
-where
-\begin{align}
-  c_L \equiv \p{-1}^{L+\lfloor L/2\rfloor}
-  \sqrt{2L+1}\, \f{L!}{\p{2L+1}!},
-  &&
-  n_L \equiv \prod_{\ell=-L}^L\p{n-\ell}.
 \end{align}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Weak SOC in the fully symmetric manifold}
+
+Here we derive the effective Hamiltonian induced on the fully
+symmetric manifold $\M_\FS$ by the single-particle SOC Hamiltonian
+$H_{\t{lat}}^{(\phi)}$ in \eqref{eq:H_lat_SOC_B} in the weak SOC limit
+$\abs{\phi}\ll1$.  For brevity, we work in units with the lattice
+spacing $a=1$ throughout this section.  Perturbations $V$ to the
+interaction Hamiltonian $H_{\t{int}}$ induce the following effective
+Hamiltonians on the fully symmetric manifold
+$\M_\FS$\cite{bravyi2011schrieffer, perlin2019effective}:
+\begin{align}
+  H_{\t{eff}}^{(1)} = \P_0 V \P_0,
+  &&
+  H_{\t{eff}}^{(2)} = -\P_0 V \E V \P_0,
+  &&
+  \P_0 = \sum_{E_{\varepsilon}=0} \op{E_{\varepsilon}},
+  &&
+  \E \equiv \sum_{E_\varepsilon>0} \f{\op{E_\varepsilon}}{E_\varepsilon},
+  \label{eq:pert}
+\end{align}
+where $\ket{E_\varepsilon}$ is an eigenstate of the interaction
+Hamiltonian $H_{\t{int}}$ with energy $E_\varepsilon$ relative to the
+interaction energy of fully symmetric states
+$\ket{\psi_\FS}\in\M_\FS$.  The sum over all $E_\varepsilon>0$ is a
+sum over a complete basis of states $\set{\ket{E_\varepsilon}}$ for
+the orthogonal complement of $\M_\FS$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Effective SOC at first order}
+\label{sec:H_eff_1}
+
+The single-particle SOC Hamiltonian is
+\begin{align}
+  H_{\t{lat}}^{(\phi)}
+  = -t \sum_{q,L} w_L\p{q} A_L^{(\phi)} D_{L,0}^{(q)},
+  &&
+  w_L\p{q} \equiv
+  \begin{cases}
+    \cos q & L~\t{even} \\
+    \sin q & L~\t{odd}
+  \end{cases},
+  \label{eq:H_lat_SOC_A}
+\end{align}
+where $A_L^{(\phi)}$ is essentially a sum over Clebsch-Gordan
+coefficients, defined in \eqref{eq:A_L_phi}, and scales as
+$A_L^{(\phi)} \sim \phi^L + O\p{\phi^{L+2}}$ when $\abs{\phi}\ll1$.
+When the SOC angle $\phi$ is sufficiently small in magnitude, the
+effective Hamiltonian induced on the fully symmetric manifold $\M_\FS$
+at first order in the perturbation $H_{\t{lat}}^{(\phi)}$ is
+\begin{align}
+  H_{\t{eff}}^{(1)}
+  = \P_0 H_{\t{lat}}^{(\phi)} \P_0
+  = -t \sum_{q,L} w_L\p{q} A_L^{(\phi)} \P_0 D_{L,0}^{(q)} \P_0,
+\end{align}
+where we can freely change the index of each single-mode drive
+operator $D_{L,0}^{(q)}$ within the fully symmetric manifold to
+re-write
+\begin{align}
+  \P_0 D_{L,0}^{(q)} \P_0 = \f1N \sum_p \P_0 D_{L,0}^{(p)} \P_0
+  = \f1N \P_0 \D_{L,0} \P_0,
+\end{align}
+which implies that
+\begin{align}
+  H_{\t{eff}}^{(1)}
+  = -t \sum_L \EE_q\sp{w_L\p{q}} A_L^{(\phi)} \D_{L,0},
+  &&
+  \EE_q\sp{X_q} \equiv \f1N \sum_q X_q.
+  \label{eq:H_eff_1}
+\end{align}
+Here $\EE_q\sp{X_q}$ is an average of $X_q$ over all occupied
+quasi-momenta $q$.  Note that we have dropped the explicit projection
+operators $\P_0$ in \eqref{eq:H_eff_1} with the understanding that we
+are implicitly restricting all effective dynamics to the fully
+symmetric manifold $\M_\FS$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Effective SOC at second order}
+\label{sec:H_eff_2}
+
+In order to expand the second-order effective Hamiltonian
+$H_{\t{eff}}^{(2)}$ induced on the fully symmetric manifold $\M_\FS$
+by the SOC Hamiltonian $H_{\t{lat}}^{(\phi)}$ in
+\eqref{eq:H_lat_SOC_A} when the SOC angle $\abs{\phi}\ll1$, we note
+that $w_L\p{q}$ in \eqref{eq:H_lat_SOC_A} takes the form
+\begin{align}
+  w_L\p{q} = \f12 \sum_{s\in\set{\pm1}} \chi_{Ls} e^{isq},
+  \label{eq:w_chi}
+\end{align}
+for some appropriately defined coefficients $\chi_{Ls}$, which implies
+that
+\begin{align}
+  H_{\t{lat}}^{(\phi)}
+  = -\f{t}{2} \sum_{\substack{L\\s\in\set{\pm1}}}
+  \chi_{Ls} A_L^{(\phi)} \Z_{Ls},
+  &&
+  \Z_{Ks} \equiv \sum_q e^{isq} D_{L,0}^{(q)},
+\end{align}
+and in turn
+\begin{align}
+  H_{\t{eff}}^{(2)}
+  = -\f{t^2}{4} \sum_{\substack{J,K\\r,s\in\set{\pm1}}}
+  \chi_{Jr} \chi_{Ks} A_J^{(\phi)} A_K^{(\phi)}
+  \P_0 \Z_{Jr} \E \Z_{Ks} \P_0,
+  \label{eq:H_eff_2_start}
+\end{align}
+where $\P_0$ is a projector onto the fully symmetric manifold
+$\M_\FS$, and $\E$ is a sum over excited-state projectors with
+energetic suppression factors, defined in \eqref{eq:pert}.  The
+remainder of this section is devoted to the task of simplifying
+\eqref{eq:H_eff_2_start} to arrive at our final result in
+\eqref{eq:H_eff_2}, expressed in purely in terms of collective drive
+operators.  We will break up this task into three steps:
+\begin{enumerate*}
+\item eliminating the excited-state projectors in $\E$,
+\item evaluating the product of operators in \eqref{eq:H_eff_2_start},
+  and
+\item evaluating the sum over $r,s\in\set{\pm1}$.
+\end{enumerate*}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Mean-subtracted spin-wave operators and the elimination
+  of excited-state projectors}
+
+In order to simplify the product of operators in
+\eqref{eq:H_eff_2_start}, we note that sandwiching the ``spin-wave''
+operators $\Z_{Jr},\Z_{Ks}$ by the ground-state projector $\P_0$ and
+the excited-state projectors in $\E$, picks off the components of
+these spin wave operators that are off-diagonal with respect to the
+symmetric manifold $\M_\FS$ and its orthogonal complement.  We
+therefore define the mean-subtracted spin-wave operators
+\begin{align}
+  \tilde\Z_{Ks}
+  \equiv \sum_q \p{e^{isq} - \EE_k\sp{e^{isk}}} D_{K,0}^{(q)},
+  &&
+  \EE_k\sp{X\p{k}} \equiv \f1N \sum_k X\p{k},
+\end{align}
+for which the difference $\tilde\Z_{Ks}-\Z_{Ks}\propto\D_{K,0}$ is a
+collective operator that only couples fully symmetric states
+$\ket{\psi_\FS}\in\M_\FS$ to fully symmetric states.  In particular,
+\begin{align}
+  \P_0 \tilde\Z_{Jr} \E \tilde\Z_{Ks} \P_0
+  = \P_0 \Z_{Jr} \E \Z_{Ks} \P_0,
+\end{align}
+which allows us to freely take
+$\Z_{Jr},\Z_{Ks}\to\tilde\Z_{Jr},\tilde\Z_{Ks}$ with no consequence
+for the effective Hamiltonian $H_{\t{eff}}^{(2)}$ in
+\eqref{eq:H_eff_2_start}.  Mean-subtracted spin-wave operators
+$\tilde\Z_{Ks}$ have the convenient feature that their action on any
+fully symmetric state $\ket{\psi_{\FS}}\in\M_\FS$ yields an
+eigenvector of the interaction Hamiltonian
+$H_{\t{int}}=-u\sum_{p<q}\Pi_{pq}$.  To substantiate this claim, we
+expand
+\begin{align}
+  \sum_{p<q} \Pi_{pq} \tilde\Z_{Ks} \ket{\psi_\FS}
+  = \sum_{\substack{p<q\\k}}
+  \p{e^{isk} - \EE_\ell\sp{e^{is\ell}}}
+  \Pi_{pq} D_{K,0}^{(k)} \ket{\psi_\FS}.
+  \label{eq:Z_eig_start}
+\end{align}
+If $k\notin\set{p,q}$, then the permutation operator $\Pi_{pq}$
+commutes with $D_{K,0}^{(k)}$, and in turn acts trivially on the
+permutationally-symmetric state $\ket{\psi_\FS}$, giving us
+\begin{align}
+  \sum_{\substack{p<q\\k\notin\set{p,q}}}
+  \p{e^{isk} - \EE_\ell\sp{e^{is\ell}}}
+  \Pi_{pq} D_{K,0}^{(k)} \ket{\psi_\FS}
+  = { N-1 \choose 2 } \tilde\Z_{Ks} \ket{\psi_\FS}.
+  \label{eq:Z_eig_notin}
+\end{align}
+If $k\in\set{p,q}$, then we can resolve the identity operator
+$\1=\Pi_{pq}^\dag \Pi_{pq}$ to expand
+\begin{multline}
+  \sum_{\substack{p<q\\k\in\set{p,q}}}
+  \p{e^{isk} - \EE_\ell\sp{e^{is\ell}}}
+  \Pi_{pq} D_{K,0}^{(k)} \ket{\psi_\FS}
+  = \f12 \sum_{\substack{p\ne q\\k\in\set{p,q}}}
+  \p{e^{isk} - \EE_\ell\sp{e^{is\ell}}}
+  \Pi_{pq} D_{K,0}^{(k)} \Pi_{pq}^\dag \Pi_{pq} \ket{\psi_\FS} \\
+  = \sum_q \sum_{k\ne q} \p{e^{isk} - \EE_\ell\sp{e^{is\ell}}}
+  D_{K,0}^{(q)} \ket{\psi_\FS},
+\end{multline}
+where $\sum_k \p{e^{isk} - \EE_\ell\sp{e^{is\ell}}} = 0$, so
+\begin{align}
+  \sum_{k\ne q} \p{e^{isk} - \EE_\ell\sp{e^{is\ell}}}
+  = - \p{e^{isq} - \EE_\ell\sp{e^{is\ell}}},
+\end{align}
+and in turn
+\begin{align}
+  \sum_{\substack{p<q\\k\in\set{p,q}}}
+  \p{e^{isk} - \EE_\ell\sp{e^{is\ell}}}
+  \Pi_{pq} D_{K,0}^{(k)} \ket{\psi_\FS}
+  = -\sum_q \p{e^{isq} - \EE_\ell\sp{e^{is\ell}}}
+  D_{K,0}^{(q)} \ket{\psi_\FS}
+  = - \tilde\Z_{Ks} \ket{\psi_\FS}.
+  \label{eq:Z_eig_in}
+\end{align}
+Substituting the results in \eqref{eq:Z_eig_in} and
+\eqref{eq:Z_eig_notin} into \eqref{eq:Z_eig_start} gives us
+\begin{align}
+  \sum_{p<q} \Pi_{pq} \tilde\Z_{Ks} \ket{\psi_\FS}
+  = \sp{{ N-2 \choose 2 } - 1} \tilde\Z_{Ks} \ket{\psi_\FS}.
+\end{align}
+The image of any fully symmetric state $\ket{\psi_\FS}\in\M_\FS$ under
+the action of the mean-subtracted spin-wave operator $\tilde\Z_{Ks}$
+is thus proportional to an eigenstate of $H_{\t{int}}$ with
+interaction energy $-u\sp{{ N-1 \choose 2 } - 1}$, which is $Nu$ above
+the interaction energy of $-u{N\choose2}$ of states in the fully
+symmetric manifold, so
+\begin{align}
+  \E \tilde\Z_{Ks} \P_0
+  =  \f1{Nu} \tilde\Z_{Ks} \P_0
+  = \f1{fU} \tilde\Z_{Ks} \P_0,
+  \label{eq:proj_elim}
+\end{align}
+where $f\equiv N/L$ is the filling fraction of quasi-momentum modes.
+We have thus far simplified the effective Hamiltonian in
+\eqref{eq:H_eff_2_start} to
+\begin{align}
+  H_{\t{eff}}^{(2)}
+  = -\f{t^2}{4fU} \sum_{\substack{J,K\\r,s\in\set{\pm1}}}
+  \chi_{Jr} \chi_{Ks} A_J^{(\phi)} A_K^{(\phi)}
+  \P_0 \tilde\Z_{Jr} \tilde\Z_{Ks} \P_0,
+  \label{eq:H_eff_2_mid}
+\end{align}
+leaving us the task of evaluating the product
+$\P_0 \tilde\Z_{Jr} \tilde\Z_{Ks} \P_0$, as well as the sum over
+$r,s$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{The spin-wave operator product}
+
+To evaluate the product $\P_0 \tilde\Z_{Jr} \tilde\Z_{Ks} \P_0$ of
+mean-subtracted spin-wave operators projected by $\P_0$ onto the fully
+symmetric manifold $\M_\FS$, we expand
+\begin{align}
+  \P_0 \tilde\Z_{Jr} \tilde\Z_{Ks} \P_0
+  = \sum_{p,q} \p{e^{irp}-\EE_k\sp{e^{irk}}}
+  \p{e^{isq}-\EE_\ell\sp{e^{is\ell}}}
+  \P_0 D_{J,0}^{(p)} D_{K,0}^{(q)} \P_0.
+  \label{eq:ZZ_prod_start}
+\end{align}
+Using the permutational symmetry of the fully symmetric manifold, we
+can simplify the $p=q$ terms above to
+\begin{align}
+  \sum_p \p{e^{irp}-\EE_k\sp{e^{irk}}}
+  \p{e^{isp}-\EE_\ell\sp{e^{is\ell}}}
+  \P_0 D_{J,0}^{(p)} D_{K,0}^{(p)} \P_0
+  = N \cov_{p,q}\sp{e^{irp},e^{isq}}
+  \P_0 D_{J,0}^{(0)} D_{K,0}^{(0)} \P_0,
+\end{align}
+where
+\begin{align}
+  \cov_{p,q}\sp{X_p,Y_q}
+  \equiv \EE_p\sp{\p{X_p-\EE_k\sp{X_k}}\p{Y_p-\EE_\ell\sp{Y_\ell}}}
+  = \EE_k\sp{X_k Y_k} - \EE_p\sp{X_p} \EE_q\sp{X_q}
+  \label{eq:cov}
+\end{align}
+is the covariance between $X_p$ and $Y_q$, evaluated over all occupied
+quasi-momenta $p,q$.  Similarly, the $p\ne q$ terms in
+\eqref{eq:ZZ_prod_start} simplify to
+\begin{multline}
+  \sum_{p\ne q} \p{e^{irp}-\EE_k\sp{e^{irk}}}
+  \p{e^{isq}-\EE_\ell\sp{e^{is\ell}}}
+  \P_0 D_{J,0}^{(p)} D_{K,0}^{(q)} \P_0 \\
+  = \sum_p \p{e^{irp}-\EE_k\sp{e^{irk}}}
+  \sum_{q\ne p} \p{e^{isq}-\EE_\ell\sp{e^{is\ell}}}
+  \P_0 D_{J,0}^{(0)} D_{K,0}^{(1)} \P_0,
+\end{multline}
+where $\sum_q \p{e^{isq}-\EE_\ell\sp{e^{is\ell}}} = 0$, so
+\begin{align}
+  \sum_{q\ne p} \p{e^{isq}-\EE_\ell\sp{e^{is\ell}}}
+  = - \p{e^{isp}-\EE_\ell\sp{e^{is\ell}}},
+\end{align}
+and in turn
+\begin{align}
+  \sum_{p\ne q} \p{e^{irp}-\EE_k\sp{e^{irk}}}
+  \p{e^{isq}-\EE_\ell\sp{e^{is\ell}}}
+  \P_0 D_{J,0}^{(p)} D_{K,0}^{(q)} \P_0
+  = - N \cov_{p,q}\sp{e^{irp},e^{isq}}
+  \P_0 D_{J,0}^{(0)} D_{K,0}^{(0)} \P_0.
+\end{align}
+Thus far, we have that
+\begin{align}
+  \P_0 \tilde\Z_{Jr} \tilde\Z_{Ks} \P_0
+  = N \cov_{p,q}\sp{e^{irp},e^{isq}}
+  \p{\P_0 D_{J,0}^{(0)} D_{K,0}^{(0)} \P_0
+    - \P_0 D_{J,0}^{(0)} D_{K,0}^{(1)} \P_0}.
+  \label{eq:ZZ_prod_mid}
+\end{align}
+In order to write this result in terms of collective operators, we
+expand
+\begin{multline}
+  \P_0 \D_{J,0} \D_{K,0} \P_0
+  = \sum_p \P_0 D_{J,0}^{(p)} D_{K,0}^{(p)} \P_0
+  + \sum_{p\ne q} \P_0 D_{J,0}^{(p)} D_{K,0}^{(q)} \P_0 \\
+  = N \P_0 D_{J,0}^{(0)} D_{K,0}^{(0)} \P_0
+  + N \p{N-1} \P_0 D_{J,0}^{(0)} D_{K,0}^{(1)} \P_0,
+\end{multline}
+which implies that
+\begin{align}
+  - N \P_0 D_{J,0}^{(0)} D_{K,0}^{(1)} \P_0
+  = \f{N}{N-1} \P_0 D_{J,0}^{(0)} D_{K,0}^{(0)} \P_0
+  - \f1{N-1} \P_0 \D_{J,0} \D_{K,0} \P_0.
+\end{align}
+The last ingredient we need to re-write \eqref{eq:ZZ_prod_mid} in
+terms of collective operators is
+\begin{align}
+  N \P_0 D_{J,0}^{(0)} D_{K,0}^{(0)} \P_0
+  = N \sum_L f_{J,0;K,0}^{L,0} \P_0 D_{L,0}^{(0)} \P_0
+  = \sum_L f_{J,0;K,0}^{L,0} \P_0 \D_{L,0} \P_0,
+\end{align}
+where $f_{J,0;K,0}^{L,0}$ is a structure constant of the transition
+operator algebra, defined in \eqref{eq:trans_prod} of Appendix
+\ref{sec:trans_prod}.  Altogether, we then have that
+\begin{align}
+  N\p{\P_0 D_{J,0}^{(0)} D_{K,0}^{(0)} \P_0
+    - \P_0 D_{J,0}^{(0)} D_{K,0}^{(1)} \P_0}
+  = \f{N}{N-1} \sum_L f_{J,0;K,0}^{L,0} \P_0 \D_{L,0} \P_0
+  - \f1{N-1} \P_0 \D_{J,0} \D_{K,0} \P_0,
+  \label{eq:eff_sym_drive}
+\end{align}
+which, substituted into \eqref{eq:ZZ_prod_mid}, implies that
+\begin{align}
+  \P_0 \tilde\Z_{Jr} \tilde\Z_{Ks} \P_0
+  = -\f{\cov_{p,q}\sp{e^{irp},e^{isq}}}{N-1}
+  \p{\P_0 \D_{J,0} \D_{K,0} \P_0
+    - N \sum_L f_{J,0;K,0}^{L,0} \P_0 \D_{L,0} \P_0}.
+  \label{eq:ZZ_prod}
+\end{align}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Putting it all together}
+
+Having evaluated the product $\P_0 \tilde\Z_{Jr} \tilde\Z_{Ks} \P_0$
+of spin-wave operators in \eqref{eq:ZZ_prod}, the last ingredient
+needed to simplify the effective Hamiltonian $H_{\t{eff}}^{(2)}$ in
+\eqref{eq:H_eff_2_mid} is an evaluation of the sum over
+$r,s\in\set{\pm1}$.  We can use the implicit definition of the
+coefficients $\chi_{Ls}$ in \eqref{eq:w_chi} together with the
+bilinearity of the covariance, defined in \eqref{eq:cov}, to simplify
+\begin{align}
+  \f14 \sum_{r,s\in\set{\pm1}} \chi_{Jr} \chi_{Ks}
+  \cov_{p,q}\sp{e^{irp},e^{isq}}
+  = \cov_{p,q}\sp{w_J\p{p},w_K\p{q}}.
+  \label{eq:sign_sum}
+\end{align}
+Substituting the operator product in \eqref{eq:ZZ_prod} into the
+effective Hamiltonian $H_{\t{eff}}^{(2)}$ in \eqref{eq:H_eff_2_mid}
+and evaluating the sum over $r,s\in\set{\pm1}$ according to
+\eqref{eq:sign_sum} gives us
+\begin{align}
+  H_{\t{eff}}^{(2)}
+  = \f{t^2}{\p{N-1}fU} \sum_{J,K} \cov_{p,q}\sp{w_J\p{pa},w_K\p{qa}}
+  A_J^{(\phi)} A_K^{(\phi)}
+  \p{\D_{J,0} \D_{K,0} - N \sum_L f_{J,0;K,0}^{L,0} \D_{L,0}},
+  \label{eq:H_eff_2}
+\end{align}
+where we have recovered factors of the lattice spacing $a$ as
+appropriate, and we have dropped the explicit projection operators
+$\P_0$ with the understanding that we are implicitly restricting all
+effective dynamics to the fully symmetric manifold $\M_\FS$.  Note
+that as the spatial filling fraction $f\equiv N/L\to1$, such that sums
+over occupied quasi-momenta $q$ become sums over nearly all angles
+$q\in\ZZ_N\times2\pi/N$, when both $J,K$ are even or odd
+(i.e.~$J=K~\t{mod}~2$) the covariance in \eqref{eq:H_eff_2} becomes
+$\cov_{p,q}\sp{\sin\p{p},\sin\p{p}}\to1/2$ or
+$\cov_{p,q}\sp{\cos\p{p},\cos\p{p}}\to1/2$, whereas when one of $J,K$
+is even and the other is odd (i.e.~$J\ne K~\t{mod}~2$), this
+covariance becomes $\cov_{p,q}\sp{\cos\p{p},\sin\p{q}}\to0$.  In the
+case of SU(2), the effective Hamiltonian $H_{\t{eff}}^{(2)}$ in
+\eqref{eq:H_eff_2} reduces to the one-axis twisting Hamiltonian
+derived in Ref.~\cite{he2019engineering}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Extracting dependence on the SOC angle $\phi$}
+
+The effective Hamiltonians in \eqref{eq:H_eff_1} and
+\eqref{eq:H_eff_2} are respectively first and second order in the SOC
+Hamiltonian $H_{\t{lat}}^{(\phi)}$.  Here we extract the leading and
+next-to-leading order dependence of these effective Hamiltonians on
+the SOC angle $\phi$.  Mindful of the expansion for the coefficients
+$A_L^{(\phi)} \sim \phi^L + O\p{\phi^{L+2}}$ in
+\eqref{eq:A_L_phi_small}, we define the re-scaled collective drive
+operators
+\begin{align}
+  \tilde\D_{LM} \equiv c_L \sqrt{n_L} \D_{LM},
+\end{align}
+in terms of which
+\begin{align}
+  H_{\t{eff}}^{(1)}
+  &= -\phi t \p{\EE_p\sp{\sin\p{pa}} \tilde\D_{1,0}
+    + \EE_q\sp{\cos\p{qa}} \tilde\D_{2,0}}
+  + O\p{\phi^3},
+  \label{eq:H_eff_1_phi} \\
+  H_{\t{eff}}^{(2)}
+  &= \f{\p{\phi t}^2}{\p{N-1}fU} \var_q\sp{\sin\p{qa}}
+  \p{\tilde\D_{1,0}^2 + 2 N \tilde\D_{2,0}}
+  + O\p{\phi^3},
+  \label{eq:H_eff_2_phi}
+\end{align}
+where
+\begin{align}
+  \var_q\sp{X_q} \equiv \cov_{p,q}\sp{X_p,X_q}
+  = \EE_p\sp{X_p^2} - \EE_q\sp{X_q}^2
+\end{align}
+is the variance of $X_q$ over all occupied quasi-momenta $q$, and we
+have neglected scalar terms $\sim\D_{0,0}\propto\1$ that have no
+physical consequence.  The effective Hamiltonians in
+\eqref{eq:H_eff_1_phi} and \eqref{eq:H_eff_2_phi} reduce to those
+derived in Ref.~\cite{he2019engineering} in the case of SU(2).
+
 \bibliography{multilevel_spin_notes.bib}

### 2019-10-25 collective_sun/multilevel_spin_notes.tex
added content to the section in the main text about perturbative SOC introduced a scaling factor for drive operators

@@
 \end{align}
-where $u\equiv U/L$ is the two-body on-site interaction energy $U$
-divided by the number of lattice sites $L$.  Defining the spin
-operators $S_{\mu\nu}^{(p)}\equiv c_{p\mu}^\dag c_{p\nu}$ for each
+where $u\equiv fU/N$ with $U$ the two-body on-site interaction energy,
+$f$ the filling fraction of lattice sites, and $N$ the total number of
+fermions.  Defining the spin operators
+$S_{\mu\nu}^{(p)}\equiv c_{p\mu}^\dag c_{p\nu}$ for each
 quasi-momentum $p$, we can alternately write
@@
 field and a classical spin-polarized driving laser; loosely speaking,
-one can identify $\p{D_{1,0},D_{1,1},D_{1,-1}}\sim\p{S_\z,S_\x,S_\y}$,
-where $S_\alpha$ is a collective SU(2) spin-$\alpha$ operator for
-$\p{n-1}$ 2-level spins\cite{perlin2019shorttime}.  More concretely,
-we can embed the Hilbert space of a single $n$-level spin into the
+one can associate
+$\p{D_{1,0},D_{1,1},D_{1,-1}}\sim\p{S_\z,S_\x,S_\y}$, where $S_\alpha$
+is a collective SU(2) spin-$\alpha$ operator for $\p{n-1}$ 2-level
+spins\cite{perlin2019shorttime}.  More concretely, we can embed the
+Hilbert space of a single $n$-level spin into the
 permutationally-symmetric (Dicke) manifold of states for $\p{n-1}$
@@
   \propto S_+^{I+\mu} \ket\dn^{\otimes\p{n-1}},
+  \label{eq:collective_embedding}
 \end{align}
@@
 \begin{align}
-  \p{D_{1,0},D_{1,1},D_{1,-1}}
-  \to \p{S_\z,S_\x,S_\y}
-  \times \f{2\sqrt{3}}{\sqrt{\p{n-1}n\p{n+1}}}.
+  \xi_1 \times \p{D_{1,0}, D_{1,1}, D_{1,-1}}
+  \to \p{S_\z, S_\x, S_\y},
   \label{eq:spin_ops}
 \end{align}
-Furthermore, if we {\it define} the operators $S_\alpha$ for a single
-$n$-level spin via the correspondence in \eqref{eq:spin_ops}, then the
-familiar SU(2) transformation rule
-\begin{align}
-  e^{i\phi S_\z} S_\x e^{-i\phi S_\z} = S_\x^{(\phi)},
-  &&
-  S_\x^{(\phi)} \equiv \cos\phi S_\x - \sin\phi S_\y,
-\end{align}
-generalizes to
-\begin{align}
-  e^{i\phi S_\z} D_{LM} e^{-i\phi S_\z} = D_{LM}^{(\phi)},
-  &&
-  D_{LM}^{(\phi)} \equiv \cos\p{M\phi} D_{LM} - \sin\p{M\phi} D_{L,-M}.
-\end{align}
+where
+\begin{align}
+  \xi_L
+  \equiv \sqrt{2L+1}\, \f{L!}{\p{2L+1}!}
+  \sp{\prod_{\ell=-L}^L\p{n+\ell}}^{1/2},
+  \label{eq:scale_fac}
+\end{align}
+is a scale factor that will appear frequently in the theory of drive
+operators.
 
@@
 $\M_{\FS}$ is the set of {\it fully symmetric} (FS) states that have
-eigenvalue $1$ with respect to all pairwise permutation operators
+eigenvalue 1 with respect to all pairwise permutation operators
 $\Pi_{pq}$, and thereby also all $\Pi_\sigma\in\Pi$.  The fully
@@
 
-A driving laser addressing spin-$1/2$ fermions on a 1-D lattice will
+A driving laser addressing spin-1/2 fermions on a 1-D lattice will
 induce the Hamiltonian
@@
   \left. H_{\t{drive}}^{(\phi)} \right|_{n=2}
-  = \Omega \sum_j
+  = \f12 \Omega \sum_j
   \p{e^{i\phi j} c_{j,\up}^\dag c_{j,\dn}
@@
   H_{\t{drive}}^{(\phi)}
-  = -\Omega \sum_j \p{e^{i\phi j} T_{1,1}^{(j)}
+  = -\f12 \Omega \sum_j \p{e^{i\phi j} T_{1,1}^{(j)}
     + e^{-i\phi j} {T_{1,1}^{(j)}}^\dag}
-  = \Omega \D_{1,1}^{(\phi)},
+  = \tilde \D_{1,1}^{(\phi)},
+  &&
+  \tilde\Omega \equiv \Omega \xi_1
   \label{eq:H_drive}
@@
  \begin{align}
-  H_{\t{drive}}^{(\phi)} \to H_{\t{drive}} = \Omega \D_{1,1},
+   H_{\t{drive}}^{(\phi)} \to H_{\t{drive}} = \tilde\Omega \D_{1,1},
    &&
@@
 \begin{align}
-  A_L^{(\phi)} = c_L \sqrt{n_L}\, \phi^L + O\p{\phi^{L+2}},
-  &&
-  c_L \equiv \p{-1}^{L+\lfloor L/2\rfloor}
-  \f{\sqrt{2L+1}\,L!}{\p{2L+1}!},
-  &&
-  n_L \equiv \prod_{\ell=-L}^L\p{n+\ell}.
+  A_L^{(\phi)}
+  = \p{-1}^{L+\lfloor L/2\rfloor} \xi_L \phi^L + O\p{\phi^{L+2}},
   \label{eq:A_L_phi_small}
 \end{align}
-
-\vspace{3cm}
+where $\xi_L$ is a scale factor defined in \eqref{eq:scale_fac}.  A
+perturbative treatment of SOC through second order in the SOC angle
+$\phi$ yields the following effective Hamiltonian in the fully
+symmetric manifold $\M_\FS$ (see Appendix \ref{sec:SOC_pert}):
+\begin{align}
+  H_{\t{eff}} = H_{\t{eff}}^{(1)} + H_{\t{eff}}^{(2)},
+\end{align}
+\begin{align}
+  H_{\t{eff}}^{(1)}
+  &= \phi t \, \EE_p\sp{\sin\p{pa}} \xi_1 \D_{1,0}
+  + \phi^2 t \, \EE_q\sp{\cos\p{qa}} \xi_2 \D_{2,0}
+  + O\p{\phi^3},
+  \label{eq:H_eff_1_phi} \\
+  H_{\t{eff}}^{(2)}
+  &= \f{\phi^2 t^2}{\p{N-1}fU} \var_p\sp{\sin\p{pa}}
+  \p{\xi_1^2 \D_{1,0}^2 - 2 N \xi_2 \D_{2,0}}
+  + O\p{\phi^3},
+  \label{eq:H_eff_2_phi}
+\end{align}
+where
+\begin{align}
+  \EE_p\sp{X_p} \equiv \f1N \sum_p X_p,
+  &&
+  \var_p\sp{X_p} \equiv \EE_p\sp{\p{X_p-\EE_q\sp{X_q}}}
+  = \EE_p\sp{X_p^2}-\EE_q\sp{X_q}^2,
+\end{align}
+are the mean and variance of $X_p$ over all occupied quasi-momenta
+$p$.  In the case of SU(2), the effective Hamiltonians in
+\eqref{eq:H_eff_1_phi} and \eqref{eq:H_eff_2_phi} reduce to the
+spin-squeezing Hamiltonians derived in Ref.~\cite{he2019engineering}.
+To facilitate comparison, we note that in the case of SU(2) the drive
+operator $\D_{1,0}$ is precisely the collective spin-$z$ operator
+$S_\z$ that appears in Ref.~\cite{he2019engineering}.  Furthermore,
+drive operators $D_{LM}$ vanish when $L\ge n$, so in the case of SU(2)
+the operator $\D_{2,0}$ has no contribution to the effective
+Hamiltonians in \eqref{eq:H_eff_2_phi} and \eqref{eq:H_eff_2_phi}.
 
@@
 amplitude; $\phi_m$ is the relative phase of drive $m$ between
-adjacent lattice sites; $\op{r}{s}$ for $r,s\in\set{\up,\dn}$ is a
+adjacent lattice sites; we have added an overall factor of $\xi_1$ to
+simplify later expressions; $\op{r}{s}$ for $r,s\in\set{\up,\dn}$ is a
 pseudo-spin operator for the auxiliary degree of freedom that is
@@
   &= -\sum_j\sum_m \f{\Omega_m^2}{\Delta}
-  {T_{1,m}^{(j)}}^\dag T_{1,m}^{(j)}, & \label{eq:raman_0} \\
+  {T_{1,m}^{(j)}}^\dag T_{1,m}^{(j)},
+  \label{eq:raman_0} \\
   H_{\t{Raman},1}^{(\phi)}
@@
   = \sum_j \sum_{\substack{L\le2\\\abs{M}\le L}}
-  \Omega_{LM}^{(\phi,j)} D_{LM}^{(j)},
+  \Omega_{LM}^{(\phi,j)} \xi_L D_{LM}^{(j)},
   \label{eq:raman_drive}
@@
   \left. H_{\t{Raman}}^{(\phi)} \right|_{\phi_m=m\phi}
-  = \sum_{\substack{L\le2\\0\le M\le L}} \Omega_{LM} \D_{LM}^{(\phi)},
+  = \sum_{\substack{L\le2\\0\le M\le L}}
+  \Omega_{LM} \xi_L \D_{LM}^{(\phi)},
   \label{eq:drive_raman_matched}
@@
   \to H_{\t{Raman}}
-  = \sum_{\substack{L\le2\\0\le M\le L}} \Omega_{LM} \D_{LM}.
+  = \sum_{\substack{L\le2\\0\le M\le L}} \Omega_{LM} \xi_L \D_{LM}.
 \end{align}
@@
 $\B\p{\H_n}$ of linear operators on the Hilbert space $\H_n$ of an
-$n$-level system, up to appropriate re-scaling factors $\N_X/\N_Y$.
+$n$-level system, up to appropriate rescaling factors $\N_X/\N_Y$.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Raman drive coefficients}
-\label{sec:drive_raman_coeff}
-
-Here we provide the coefficients $\Omega_{LM}^{(\phi,j)}$ of the
-effective Raman driving Hamiltonian
-\begin{align}
-  H_{\t{Raman}}^{(\phi)}
-  = \sum_j \sum_{\substack{L\in\set{0,1,2}\\\abs{M}\le L}}
-  \Omega_{LM}^{(\phi,j)} D_{LM}^{(j)}
-\end{align}
-considered in Section \ref{sec:drive_raman}.  These coefficients are
-computed using symmetry properties and structure constants of the
-transition operator algebra (see Appendix \ref{sec:trans_prod}).  For
-clarity, if any coefficient $\Omega_{LM}^{(\phi,j)}$ is independent of
-$\phi$, we drop the explicit dependence on $\phi$, i.e.~taking
-$\Omega_{LM}^{(\phi,j)}\to\Omega_{LM}$.
-\begin{align}
-  \Omega_{0,0} \equiv -\f1{\sqrt{n}}\,
-  \sum_m\f{\Omega_m^2}{\Delta},
-  &&
-  \Omega_{1,0} \equiv \sqrt{\f{3}{n_1}}\,
-  \f{\Omega_+^2 - \Omega_-^2}{\Delta},
-  \label{eq:O_01_0}
-\end{align}
-\begin{align}
-  \Omega_{2,0} \equiv \f1{\sqrt{5}}\, \f{\sqrt{n_2}}{n_1}\,
-  \f{\Omega_+ + \Omega_- - 2\Omega_0}{\Delta},
-  \label{eq:O_2_0}
-\end{align}
-\begin{align}
-  \Omega_{L,1}^{(\phi,j)}
-  &\equiv \sum_{s\in\set{\pm1}} \omega_{Ls} \cos\p{\phi_s j},
-  &
-  \Omega_{L,-1}^{(\phi,j)}
-  &\equiv -\sum_{s\in\set{\pm1}} s \omega_{Ls} \sin\p{\phi_s j}, \\
-  \Omega_{2,2}^{(\phi,j)}
-  &\equiv \omega_{2,2} \cos\p{\sp{\phi_+-\phi_-}j},
-  &
-  \Omega_{2,-2}^{(\phi,j)}
-  &\equiv -\omega_{2,2} \sin\p{\sp{\phi_+-\phi_-}j}, \\
-  \omega_{1\pm}
-  &\equiv \sqrt{\f{6}{n_1}}\, \f{\Omega_0 \Omega_\pm}{\Delta},
-  &
-  \omega_{2\pm}
-  &\equiv \pm \sqrt{\f65}\, \f{\sqrt{n_2}}{n_1}\,
-  \f{\Omega_0\Omega_\pm}{\Delta}, \\
-  \omega_{2,2}
-  &\equiv 2\sqrt{\f35}\, \f{\sqrt{n_2}}{n_1}\,
-  \f{\Omega_+\Omega_-}{\Delta},
-  &
-  n_L
-  &\equiv \prod_{\ell=-L}^L \p{n+\ell}.
-\end{align}
-Here $\Omega_m$, $\phi_m$, and $\Delta$ are respectively the (real)
-driving amplitudes, phases, and detuning that appear in the ``bare''
-Raman drive in \eqref{eq:raman_bare}, and by choice of gauge we set
-$\phi_-=0$, or equivalently re-define $\phi_m-\phi_0\to\phi_m$.  If
-the driving lasers are phase-matched with $\phi_m=m\phi$, such that a
-nuclear spin transition $\mu\to\mu+m$ on site $j$ imprints the phase
-$m\phi j$, then the drive $H_{\t{Raman}}^{(\phi)}$ takes the form
-\begin{align}
-  \left. H_{\t{Raman}}^{(\phi)} \right|_{\phi_m=m\phi}
-  = \sum_{\substack{L\le2\\0\le M\le L}} \Omega_{LM} \D_{LM}^{(\phi)},
-\end{align}
-with inhomogeneously rotated collective drive operators
-\begin{align}
-  \D_{LM}^{(\phi)}
-  \equiv \sum_j \sp{\cos\p{M\phi j} D_{LM}^{(j)}
-    - \sin\p{M\phi j} D_{L,-M}^{(j)}},
-\end{align}
-and uniform coefficients given by \eqref{eq:O_01_0}, \eqref{eq:O_2_0},
-and
-\begin{align}
-  \Omega_{1,-1} = \Omega_{2,-1} = \Omega_{2,-2} \equiv 0,
-  &&
-  \Omega_{L,1} \equiv \sum_{s\in\set{\pm1}} \omega_{Ls},
-  &&
-  \Omega_{2,2} \equiv \omega_{2,2}.
-\end{align}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Weak SOC in the fully symmetric manifold}
+\label{sec:SOC_pert}
 
@@
 \end{align}
-where $f\equiv N/L$ is the filling fraction of quasi-momentum modes.
-We have thus far simplified the effective Hamiltonian in
-\eqref{eq:H_eff_2_start} to
+where $f$ is the filling fraction of lattice sites.  We have thus far
+simplified the effective Hamiltonian in \eqref{eq:H_eff_2_start} to
 \begin{align}
@@
 effective dynamics to the fully symmetric manifold $\M_\FS$.  Note
-that as the spatial filling fraction $f\equiv N/L\to1$, such that sums
-over occupied quasi-momenta $q$ become sums over nearly all angles
+that as the spatial filling fraction $f\to1$, such that sums over
+occupied quasi-momenta $q$ become sums over nearly all angles
 $q\in\ZZ_N\times2\pi/N$, when both $J,K$ are even or odd
@@
 next-to-leading order dependence of these effective Hamiltonians on
-the SOC angle $\phi$.  Mindful of the expansion for the coefficients
-$A_L^{(\phi)} \sim \phi^L + O\p{\phi^{L+2}}$ in
-\eqref{eq:A_L_phi_small}, we define the re-scaled collective drive
-operators
-\begin{align}
-  \tilde\D_{LM} \equiv c_L \sqrt{n_L} \D_{LM},
-\end{align}
-in terms of which
+the SOC angle $\phi$ using the expansion
 \begin{align}
   H_{\t{eff}}^{(1)}
-  &= -\phi t \p{\EE_p\sp{\sin\p{pa}} \tilde\D_{1,0}
-    + \EE_q\sp{\cos\p{qa}} \tilde\D_{2,0}}
+  &= \phi t \, \EE_p\sp{\sin\p{pa}} \xi_1 \D_{1,0}
+  + \phi^2 t \, \EE_q\sp{\cos\p{qa}} \xi_2 \D_{2,0}
   + O\p{\phi^3},
-  \label{eq:H_eff_1_phi} \\
+  \label{eq:H_eff_1_phi_apndx} \\
   H_{\t{eff}}^{(2)}
-  &= \f{\p{\phi t}^2}{\p{N-1}fU} \var_q\sp{\sin\p{qa}}
-  \p{\tilde\D_{1,0}^2 + 2 N \tilde\D_{2,0}}
+  &= \f{\phi^2 t^2}{\p{N-1}fU} \var_p\sp{\sin\p{pa}}
+  \p{\xi_1^2 \D_{1,0}^2 - 2 N \xi_2 \D_{2,0}}
   + O\p{\phi^3},
-  \label{eq:H_eff_2_phi}
-\end{align}
-where
+  \label{eq:H_eff_2_phi_apndx}
+\end{align}
+where the factors $\xi_L$ are defined in \eqref{eq:scale_fac},
 \begin{align}
   \var_q\sp{X_q} \equiv \cov_{p,q}\sp{X_p,X_q}
-  = \EE_p\sp{X_p^2} - \EE_q\sp{X_q}^2
+  = \EE_p\sp{\p{X_p-\EE_q\sp{X_q}}^2}
 \end{align}
@@
 have neglected scalar terms $\sim\D_{0,0}\propto\1$ that have no
-physical consequence.  The effective Hamiltonians in
-\eqref{eq:H_eff_1_phi} and \eqref{eq:H_eff_2_phi} reduce to those
-derived in Ref.~\cite{he2019engineering} in the case of SU(2).
+physical consequence.  Note that in the case of SU(2), the drive
+operator $\tilde\D_{1,0}$ is precisely the collective spin-$z$
+operator $S_\z$ that appears in Ref.~\cite{he2019engineering}.
+Furthermore, drive operators $D_{LM}$ vanish when $L\ge n$, so in the
+case of SU(2) the operator $\tilde\D_{2,0}$ has no contribution to the
+effective Hamiltonians in \eqref{eq:H_eff_2_phi_apndx} and
+\eqref{eq:H_eff_2_phi_apndx}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Raman drive coefficients}
+\label{sec:drive_raman_coeff}
+
+Here we provide the coefficients $\Omega_{LM}^{(\phi,j)}$ of the
+effective Raman driving Hamiltonian
+\begin{align}
+  H_{\t{Raman}}^{(\phi)}
+  = \sum_j \sum_{\substack{L\in\set{0,1,2}\\\abs{M}\le L}}
+  \Omega_{LM}^{(\phi,j)} D_{LM}^{(j)}
+  \label{eq:raman_drive_apndx}
+\end{align}
+considered in Section \ref{sec:drive_raman}.  These coefficients are
+computed using the structure constants of the transition operator
+algebra provided in Appendix \ref{sec:trans_prod}.  For clarity, if
+any coefficient $\Omega_{LM}^{(\phi,j)}$ is independent of $\phi$, we
+suppress the explicit dependence on $\phi$, i.e.~taking
+$\Omega_{LM}^{(\phi,j)}\to\Omega_{LM}$.  Defining
+\begin{align}
+  \tilde\Omega_m \equiv \xi_1^{-1} \Omega_m,
+  &&
+  \tilde\Omega_{LM} \equiv \xi_L^{-1} \Omega_{LM},
+\end{align}
+where the scale factors $\xi_L$ are defined in \eqref{eq:scale_fac},
+the coefficients of the effective Raman drive in
+\eqref{eq:raman_drive_apndx} are determined by
+\begin{align}
+  \Delta \tilde\Omega_{0,0}
+  \equiv -\f{\xi_1^2}{\xi_2^2} \sum_m \tilde\Omega_m^2,
+  &&
+  \Delta \tilde\Omega_{1,0}
+  \equiv \f12 \p{\tilde\Omega_+^2 - \tilde\Omega_-^2},
+  &&
+  \Delta \tilde\Omega_{2,0}
+  \equiv \tilde\Omega_+^2 + \tilde\Omega_-^2 - 2\tilde\Omega_0^2,
+  \label{eq:O_X0}
+\end{align}
+\begin{align}
+  \tilde\Omega_{L,1}^{(\phi,j)}
+  &\equiv \sum_{s\in\set{\pm1}} \omega_{Ls} \cos\p{\phi_s j},
+  &
+  \tilde\Omega_{L,-1}^{(\phi,j)}
+  &\equiv -\sum_{s\in\set{\pm1}} s \omega_{Ls} \sin\p{\phi_s j}, \\
+  \tilde\Omega_{2,2}^{(\phi,j)}
+  &\equiv \omega_{2,2} \cos\p{\sp{\phi_+-\phi_-}j},
+  &
+  \tilde\Omega_{2,-2}^{(\phi,j)}
+  &\equiv -\omega_{2,2} \sin\p{\sp{\phi_+-\phi_-}j},
+\end{align}
+\begin{align}
+  \Delta \omega_{1\pm}
+  \equiv \f1{\sqrt{2}} \, \tilde\Omega_0 \tilde\Omega_\pm,
+  &&
+  \Delta \omega_{2\pm}
+  \equiv \pm \sqrt{6} \, \tilde\Omega_0\tilde\Omega_\pm,
+  &&
+  \Delta \omega_{2,2}
+  \equiv 2\sqrt{3} \, \tilde\Omega_+\tilde\Omega_-.
+\end{align}
+Here $\tilde\Omega_m$, $\phi_m$, and $\Delta$ are respectively the
+(real) driving amplitudes, phases, and detuning that appear in the
+``bare'' Raman drive in \eqref{eq:raman_bare}, and by choice of gauge
+we set $\phi_-=0$, or equivalently re-define $\phi_m-\phi_0\to\phi_m$.
+If the driving lasers are phase-matched with $\phi_m=m\phi$, such that
+a nuclear spin transition $\mu\to\mu+m$ on site $j$ imprints the phase
+$m\phi j$, then the drive $H_{\t{Raman}}^{(\phi)}$ takes the form
+\begin{align}
+  \left. H_{\t{Raman}}^{(\phi)} \right|_{\phi_m=m\phi}
+  = \sum_{\substack{L\le2\\0\le M\le L}} \Omega_{LM} \D_{LM}^{(\phi)},
+\end{align}
+with inhomogeneously rotated collective drive operators
+\begin{align}
+  \D_{LM}^{(\phi)}
+  \equiv \sum_j \sp{\cos\p{M\phi j} D_{LM}^{(j)}
+    - \sin\p{M\phi j} D_{L,-M}^{(j)}},
+\end{align}
+and uniform coefficients determined by \eqref{eq:O_X0} and
+\begin{align}
+  \tilde\Omega_{1,-1} = \tilde\Omega_{2,-1}
+  = \tilde\Omega_{2,-2} \equiv 0,
+  &&
+  \tilde\Omega_{L,1} \equiv \sum_{s\in\set{\pm1}} \omega_{Ls},
+  &&
+  \tilde\Omega_{2,2} \equiv \omega_{2,2}.
+\end{align}
 

### 2019-10-25 collective_sun/multilevel_spin_notes.tex
some cleanup in math/conventions

@@
   H_{\t{drive}}^{(\phi)}
-  = -\f12 \Omega \sum_j \p{e^{i\phi j} T_{1,1}^{(j)}
-    + e^{-i\phi j} {T_{1,1}^{(j)}}^\dag}
-  = \tilde \D_{1,1}^{(\phi)},
-  &&
-  \tilde\Omega \equiv \Omega \xi_1
+  = -\f12 \tilde\Omega \sum_j \p{e^{i\phi j} T_{1,1}^{(j)}
+    + e^{-i\phi j} T_{1,1}^{(j)\dag}}
+  = \Omega \D_{1,1}^{(\phi)},
+  &&
+  \Omega \equiv \tilde \Omega \xi_1,
   \label{eq:H_drive}
 \end{align}
-where
+where we define the re-scaled driving amplitude $\tilde\Omega$ for
+convenience, and
 \begin{align}
@@
  \begin{align}
-   H_{\t{drive}}^{(\phi)} \to H_{\t{drive}} = \tilde\Omega \D_{1,1},
+   H_{\t{drive}}^{(\phi)} \to H_{\t{drive}} = \Omega \D_{1,1},
    &&
@@
   &= -\sum_j\sum_m \f{\Omega_m^2}{\Delta}
-  {T_{1,m}^{(j)}}^\dag T_{1,m}^{(j)},
+  T_{1,m}^{(j)\dag} T_{1,m}^{(j)},
   \label{eq:raman_0} \\
@@
   \f{\Omega_0\Omega_s}{\Delta} e^{i\p{\phi_s-\phi_0} j}
-  {T_{1,0}^{(j)}}^\dag T_{1,s}^{(j)} + \t{h.c.},
+  T_{1,0}^{(j)\dag} T_{1,s}^{(j)} + \t{h.c.},
   \\
@@
   \f{\Omega_+\Omega_-}{\Delta} e^{i\p{\phi_s-\phi_{-s}}j}
-  {T_{1,-s}^{(j)}}^\dag T_{1,s}^{(j)}.
+  T_{1,-s}^{(j)\dag} T_{1,s}^{(j)}.
   \label{eq:raman_2}
@@
   = \sum_j \sum_{\substack{L\le2\\\abs{M}\le L}}
-  \Omega_{LM}^{(\phi,j)} \xi_L D_{LM}^{(j)},
+  \Omega_{LM}^{(\phi,j)} D_{LM}^{(j)},
   \label{eq:raman_drive}
@@
   = \sum_{\substack{L\le2\\0\le M\le L}}
-  \Omega_{LM} \xi_L \D_{LM}^{(\phi)},
+  \Omega_{LM} \D_{LM}^{(\phi)},
   \label{eq:drive_raman_matched}
@@
   \to H_{\t{Raman}}
-  = \sum_{\substack{L\le2\\0\le M\le L}} \Omega_{LM} \xi_L \D_{LM}.
+  = \sum_{\substack{L\le2\\0\le M\le L}} \Omega_{LM} \D_{LM}.
 \end{align}
@@
 \begin{align}
-  H = \sum_{p,q,j} X_j^{(p)} {X_j^{(q)}}^\dag,
+  H = \sum_{p,q,j} X_j^{(p)} X_j^{(q)\dag},
 \end{align}
@@
 suppress the explicit dependence on $\phi$, i.e.~taking
-$\Omega_{LM}^{(\phi,j)}\to\Omega_{LM}$.  Defining
+$\Omega_{LM}^{(\phi,j)}\to\Omega_{LM}$.  Defining the re-scaled
+driving amplitudes
 \begin{align}
@@
   \Delta \tilde\Omega_{0,0}
-  \equiv -\f{\xi_1^2}{\xi_2^2} \sum_m \tilde\Omega_m^2,
+  \equiv -\f{\xi_1^2}{\xi_0^2} \sum_m \tilde\Omega_m^2,
   &&
@@
   \tilde\Omega_{L,1}^{(\phi,j)}
-  &\equiv \sum_{s\in\set{\pm1}} \omega_{Ls} \cos\p{\phi_s j},
+  &\equiv \sum_{s\in\set{\pm1}}
+  \omega_{Ls} \cos\p{\sp{\phi_s-\phi_0}j},
   &
   \tilde\Omega_{L,-1}^{(\phi,j)}
-  &\equiv -\sum_{s\in\set{\pm1}} s \omega_{Ls} \sin\p{\phi_s j}, \\
+  &\equiv -\sum_{s\in\set{\pm1}}
+  s \omega_{Ls} \sin\p{\sp{\phi_s-\phi_0}j}, \\
   \tilde\Omega_{2,2}^{(\phi,j)}
@@
 (real) driving amplitudes, phases, and detuning that appear in the
-``bare'' Raman drive in \eqref{eq:raman_bare}, and by choice of gauge
-we set $\phi_-=0$, or equivalently re-define $\phi_m-\phi_0\to\phi_m$.
-If the driving lasers are phase-matched with $\phi_m=m\phi$, such that
-a nuclear spin transition $\mu\to\mu+m$ on site $j$ imprints the phase
-$m\phi j$, then the drive $H_{\t{Raman}}^{(\phi)}$ takes the form
+``bare'' Raman drive in \eqref{eq:raman_bare}.  If the driving lasers
+are phase-matched with $\phi_m=m\phi$, such that a nuclear spin
+transition $\mu\to\mu+m$ on site $j$ imprints the phase $m\phi j$,
+then the drive $H_{\t{Raman}}^{(\phi)}$ takes the form
 \begin{align}

### 2019-10-25 collective_sun/multilevel_spin_notes.tex
minor changes to math and text

@@
 is simply a sum over all permutations $\Pi_{pq}$ of subsystems $p$ and
-$q$; and
+$q$; and the second sum,
 \begin{align}
@@
   = n \sum_p \1^{(p)}
-  = n N
+  = n N,
   \label{eq:spin_const}
@@
 eigenvalue 1 with respect to all pairwise permutation operators
-$\Pi_{pq}$, and thereby also all $\Pi_\sigma\in\Pi$.  The fully
-symmetric manifold $\M_\FS$ is spanned by a basis $\set{\ket{\v m}}$
-of fully symmetric states that can be uniquely identified by a list of
-integers $\v m\equiv\p{m_0,m_1,\cdots,m_{n-1}}\in\ZZ_{N+1}^n$, where
-$m_s$ indicates the total occupation number of subsystem state $s$,
-and $\sum_sm_s=N$.  In the case of $n=2$, the fully symmetric manifold
+$\Pi_{pq}$.  The fully symmetric manifold $\M_\FS$ is spanned by a
+basis $\set{\ket{\v m}}$ of fully symmetric states that can be
+uniquely identified by a list of integers
+$\v m\equiv\p{m_0,m_1,\cdots,m_{n-1}}\in\ZZ_{N+1}^n$, where $m_s$
+indicates the total occupation number of subsystem state $s$, and
+$\sum_sm_s=N$.  In the case of $n=2$, the fully symmetric manifold
 $\M_\FS$ is precisely the Dicke manifold\cite{dicke1954coherence} of
@@
 ways to assign $N$ subsystems to $n$ states, which is
-${N+n-1 \choose n-1}$, and the energy of a state
+${N+n-1 \choose n-1}$, and the energy of any fully symmetric state
 $\ket{\psi_\FS}\in\M_\FS$ with respect to the interaction Hamiltonian
@@
   H_{\t{eff}}^{(2)}
-  &= \f{\phi^2 t^2}{\p{N-1}fU} \var_p\sp{\sin\p{pa}}
+  &= \f{\phi^2 t^2}{\p{N-1}fU} \var_q\sp{\sin\p{qa}}
   \p{\xi_1^2 \D_{1,0}^2 - 2 N \xi_2 \D_{2,0}}
@@
 \begin{align}
-  \EE_p\sp{X_p} \equiv \f1N \sum_p X_p,
-  &&
-  \var_p\sp{X_p} \equiv \EE_p\sp{\p{X_p-\EE_q\sp{X_q}}}
+  \EE_q\sp{X_q} \equiv \f1N \sum_q X_q,
+  &&
+  \var_q\sp{X_q} \equiv \EE_p\sp{\p{X_p-\EE_q\sp{X_q}}}
   = \EE_p\sp{X_p^2}-\EE_q\sp{X_q}^2,
 \end{align}
-are the mean and variance of $X_p$ over all occupied quasi-momenta
-$p$.  In the case of SU(2), the effective Hamiltonians in
+are the mean and variance of $X_q$ over all occupied quasi-momenta
+$q$.  In the case of SU(2), the effective Hamiltonians in
 \eqref{eq:H_eff_1_phi} and \eqref{eq:H_eff_2_phi} reduce to the
@@
 amplitude; $\phi_m$ is the relative phase of drive $m$ between
-adjacent lattice sites; we have added an overall factor of $\xi_1$ to
-simplify later expressions; $\op{r}{s}$ for $r,s\in\set{\up,\dn}$ is a
+adjacent lattice sites; $\op{r}{s}$ for $r,s\in\set{\up,\dn}$ is a
 pseudo-spin operator for the auxiliary degree of freedom that is
@@
 lasers from the energy of an excited state; and $\1$ is the identity
-operator.  If the detuning is large, $\Delta\gg\Omega_m$, then at
-second order in perturbation theory the effective action of
+operator.  If the detuning is large, $\abs{\Delta}\gg\abs{\Omega_m}$,
+then at second order in perturbation theory the effective action of
 $H_{\t{Raman,bare}}^{(\phi)}$ within the $\ket{\dn}$-state manifold
@@
 \end{align}
-where $H_{\t{Raman},\tau}^{(\phi)}$ generates a nuclear spin
-transition $\mu\to\mu\pm\tau$:
+where $H_{\t{Raman},\tau}^{(\phi)}$ generates nuclear spin transitions
+$\mu\to\mu\pm\tau$:
 \begin{align}
@@
   &
-  \Omega_{1,1} &\propto \Omega_0 \p{\Omega_+ + \Omega_-}, \\
   \Omega_{2,0} &\propto \Omega_+^2 + \Omega_-^2 - 2\Omega_0^2,
+  \\
+  \Omega_{1,1} &\propto \Omega_0 \p{\Omega_+ + \Omega_-},
   &
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Future directions}
+\section{Possible future directions}
 
 \begin{itemize}
-\item Visualization with generalized Bloch spheres
+\item Visualization of multilevel states
   \begin{itemize}
@@
   \item Single-particle dynamics induced by drive operators
+  \item Collective states and dynamics
   \end{itemize}
-\item Classifying collective states (e.g.~eigenstates of interactions,
-  spin-wave-like states)
+\item Classifying eigenstates of SU($n$)-symmetric interactions
 \item Implications of SOC
@@
   = \EE_p\sp{\p{X_p-\EE_q\sp{X_q}}^2}
+  = \EE_p\sp{X_p^2} - \EE_q\sp{X_q}^2
 \end{align}

### 2019-10-26 collective_sun/multilevel_spin_notes.tex
write the excited-state projectors in second-order perturbation theory using eigenspace projectors, rather than eigenstates

@@
   &&
-  \P_0 = \sum_{E_{\varepsilon}=0} \op{E_{\varepsilon}},
-  &&
-  \E \equiv \sum_{E_\varepsilon>0} \f{\op{E_\varepsilon}}{E_\varepsilon},
+  \E \equiv \sum_{\Delta>0} \f{\P_\Delta}{\Delta},
   \label{eq:pert}
 \end{align}
-where $\ket{E_\varepsilon}$ is an eigenstate of the interaction
-Hamiltonian $H_{\t{int}}$ with energy $E_\varepsilon$ relative to the
-interaction energy of fully symmetric states
-$\ket{\psi_\FS}\in\M_\FS$.  The sum over all $E_\varepsilon>0$ is a
-sum over a complete basis of states $\set{\ket{E_\varepsilon}}$ for
-the orthogonal complement of $\M_\FS$.
+where $\P_\Delta$ is a projector onto the eigenspace of the
+interaction Hamiltonian $H_{\t{int}}$ corresponding to an energy
+$\Delta$ above that of fully symmetric states
+$\ket{\psi_\FS}\in\M_\FS$.
 
@@
 where $\P_0$ is a projector onto the fully symmetric manifold
-$\M_\FS$, and $\E$ is a sum over excited-state projectors with
-energetic suppression factors, defined in \eqref{eq:pert}.  The
-remainder of this section is devoted to the task of simplifying
-\eqref{eq:H_eff_2_start} to arrive at our final result in
-\eqref{eq:H_eff_2}, expressed in purely in terms of collective drive
-operators.  We will break up this task into three steps:
+$\M_\FS$, and $\E$ is a sum over projectors onto excited-state
+manifolds with corresponding energetic suppression factors, defined in
+\eqref{eq:pert}.  The remainder of this section is devoted to the task
+of simplifying \eqref{eq:H_eff_2_start} to arrive at our final result
+in \eqref{eq:H_eff_2}, expressed in purely in terms of collective
+drive operators that are implicitly restricted to the fully symmetric
+manifold $\M_\FS$.  We will break up our task into three steps:
 \begin{enumerate*}
@@
 the excited-state projectors in $\E$, picks off the components of
-these spin wave operators that are off-diagonal with respect to the
+these spin-wave operators that are off-diagonal with respect to the
 symmetric manifold $\M_\FS$ and its orthogonal complement.  We
@@
   &&
-  \EE_k\sp{X\p{k}} \equiv \f1N \sum_k X\p{k},
+  \EE_k\sp{X_k} \equiv \f1N \sum_k X_k,
 \end{align}

### 2019-10-28 collective_sun/multilevel_spin_notes.tex
simplifying some derivations

@@
   &&
-  \E \equiv \sum_{\Delta>0} \f{\P_\Delta}{\Delta},
+  \E \equiv \sum_{k>0} \f{\P_k}{\Delta_k},
   \label{eq:pert}
 \end{align}
-where $\P_\Delta$ is a projector onto the eigenspace of the
-interaction Hamiltonian $H_{\t{int}}$ corresponding to an energy
-$\Delta$ above that of fully symmetric states
+where $\P_k$ is a projector onto the $k$-th excited eigenspace of the
+interaction Hamiltonian $H_{\t{int}}$, with interaction energy
+$\Delta_k$ above that of fully symmetric states
 $\ket{\psi_\FS}\in\M_\FS$.
@@
 \end{multline}
-where $\sum_k \p{e^{isk} - \EE_\ell\sp{e^{is\ell}}} = 0$, so
+where the sum over $k\ne q$ is equal to a sum over all $k$ minus a
+term with $k=q$, i.e.
 \begin{align}
   \sum_{k\ne q} \p{e^{isk} - \EE_\ell\sp{e^{is\ell}}}
+  = \sum_k \p{e^{isk} - \EE_\ell\sp{e^{is\ell}}}
+  - \p{e^{isq} - \EE_\ell\sp{e^{is\ell}}}
   = - \p{e^{isq} - \EE_\ell\sp{e^{is\ell}}},
 \end{align}
-and in turn
+so
 \begin{align}
@@
   \Pi_{pq} D_{K,0}^{(k)} \ket{\psi_\FS}
-  = -\sum_q \p{e^{isq} - \EE_\ell\sp{e^{is\ell}}}
-  D_{K,0}^{(q)} \ket{\psi_\FS}
   = - \tilde\Z_{Ks} \ket{\psi_\FS}.
@@
 \end{multline}
-where $\sum_q \p{e^{isq}-\EE_\ell\sp{e^{is\ell}}} = 0$, so
+where the sum over $q\ne p$ is equal to a sum over all $q$ minus a
+term with $q=p$, i.e.
 \begin{align}
   \sum_{q\ne p} \p{e^{isq}-\EE_\ell\sp{e^{is\ell}}}
+  = \sum_q \p{e^{isq}-\EE_\ell\sp{e^{is\ell}}}
+  - \p{e^{isp}-\EE_\ell\sp{e^{is\ell}}}
   = - \p{e^{isp}-\EE_\ell\sp{e^{is\ell}}},
 \end{align}
-and in turn
-\begin{align}
-  \sum_{p\ne q} \p{e^{irp}-\EE_k\sp{e^{irk}}}
-  \p{e^{isq}-\EE_\ell\sp{e^{is\ell}}}
-  \P_0 D_{J,0}^{(p)} D_{K,0}^{(q)} \P_0
-  = - N \cov_{p,q}\sp{e^{irp},e^{isq}}
-  \P_0 D_{J,0}^{(0)} D_{K,0}^{(0)} \P_0.
+which implies that the overall sum over $p,q$ is
+\begin{align}
+  -\sum_p \p{e^{irp}-\EE_k\sp{e^{irk}}}
+  \p{e^{isp}-\EE_\ell\sp{e^{is\ell}}}
+  = - N \cov_{p,q}\sp{e^{irp},e^{isq}}.
 \end{align}
@@
 physical consequence.  Note that in the case of SU(2), the drive
-operator $\tilde\D_{1,0}$ is precisely the collective spin-$z$
-operator $S_\z$ that appears in Ref.~\cite{he2019engineering}.
-Furthermore, drive operators $D_{LM}$ vanish when $L\ge n$, so in the
-case of SU(2) the operator $\tilde\D_{2,0}$ has no contribution to the
-effective Hamiltonians in \eqref{eq:H_eff_2_phi_apndx} and
+operator $\D_{1,0}$ is precisely the collective spin-$z$ operator
+$S_\z$ that appears in Ref.~\cite{he2019engineering}.  Furthermore,
+drive operators $D_{LM}$ vanish when $L\ge n$, so in the case of SU(2)
+the operator $\D_{2,0}$ has no contribution to the effective
+Hamiltonians in \eqref{eq:H_eff_2_phi_apndx} and
 \eqref{eq:H_eff_2_phi_apndx}.

### 2019-10-28 collective_sun/extracting_projections.tex
adding notes on a trick to do perturbation theory within the fully symmetric manifold

\documentclass[nofootinbib,notitlepage,11pt]{revtex4-2}

%%% linking references
\usepackage{hyperref}
\hypersetup{
  breaklinks=true,
  colorlinks=true,
  linkcolor=blue,
  filecolor=magenta,
  urlcolor=cyan,
}

%%% header / footer
\usepackage{fancyhdr} % easier header and footer management
\pagestyle{fancy} % page formatting style
\fancyhf{} % clear all header and footer text
\renewcommand{\headrulewidth}{0pt} % remove horizontal line in header
\usepackage{lastpage} % for referencing last page
\cfoot{\thepage~of \pageref{LastPage}} % "x of y" page labeling

%%% symbols, notations, etc.
\usepackage{physics,braket,bm,amssymb} % physics and math
\renewcommand{\t}{\text} % text in math mode
\newcommand{\f}[2]{\dfrac{#1}{#2}} % shorthand for fractions
\newcommand{\p}[1]{\left(#1\right)} % parenthesis
\renewcommand{\sp}[1]{\left[#1\right]} % square parenthesis
\renewcommand{\set}[1]{\left\{#1\right\}} % curly parenthesis
\newcommand{\bk}{\Braket} % shorthand for braket notation
\renewcommand{\v}{\bm} % bold vectors
\newcommand{\uv}[1]{\bm{\hat{#1}}} % unit vectors
\newcommand{\av}{\vec} % arrow vectors
\renewcommand{\d}{\text{d}} % for infinitesimals
\renewcommand{\c}{\cdot} % inner product

\usepackage{dsfont} % for identity operator
\newcommand{\1}{\mathds{1}}

\newcommand{\up}{\uparrow}
\newcommand{\dn}{\downarrow}

\newcommand{\x}{\text{x}}
\newcommand{\y}{\text{y}}
\newcommand{\z}{\text{z}}

\newcommand{\B}{\mathcal{B}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathcal{E}}
\renewcommand{\H}{\mathcal{H}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\M}{\mathcal{M}}
\newcommand{\N}{\mathcal{N}}
\renewcommand{\O}{\mathcal{O}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\T}{\mathcal{T}}
\renewcommand{\S}{\mathcal{S}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\Z}{\mathcal{Z}}

\newcommand{\EE}{\mathbb{E}}
\renewcommand{\SS}{\mathbb{S}}
\newcommand{\ZZ}{\mathbb{Z}}

\newcommand{\FS}{\text{FS}}

\DeclareMathOperator{\sign}{sign}
\DeclareMathOperator{\cov}{cov}
\let\var\relax
\DeclareMathOperator{\var}{var}

\def\obra#1{\mathinner{({#1}|}}
\def\oket#1{\mathinner{|{#1})}}
\def\obk#1{\mathinner{({#1})}}
\def\oop#1#2{\oket{#1}\!\obra{#2}}

\usepackage[inline]{enumitem} % in-line lists and \setlist{} (below)
\setlist[enumerate,1]{label={(\roman*)}} % default in-line numbering
\setlist{nolistsep} % more compact spacing between environments

%%% text markup
\usepackage{color} % text color
\newcommand{\red}[1]{{\color{red} #1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\thispagestyle{fancy}

\title{Perturbation theory with collective states}%
\author{Michael A. Perlin}%
\date{\today}

\maketitle

We consider an array of $N$ multilevel spins with SU($n$)-symmetric
interactions that can be written in the form
\begin{align}
  H_{\t{int}} = -u \H,
  &&
  \H \equiv \sum_{p<q} \Pi_{pq},
  &&
  \Pi_{pq} \equiv \sum_{\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)},
\end{align}
where $S_{\mu\nu}^{(p)}\equiv\op{\mu}{\nu}_p$ flips the state of spin
$p$ to $\ket\mu$ from $\ket\nu$.  The ground-state manifold $\M_0$ of
the interaction Hamiltonian $H_{\t{int}}$ consists of fully symmetric
states with interaction energy $-{N\choose2}u$; this manifold is
gapped by an energy $Nu$ from the first excited manifold of spin-wave
states.

We wish to determine the effective dynamics induced on the
ground-state manifold $\M_0$ by weak perturbations of the form
\begin{align}
  \Z_1 \equiv \sum_p c_p Z_p,
  &&
  \Z_2 \equiv \f12 \sum_{p\ne q} c_{pq} Z_p Z_q,
\end{align}
where $Z_p$ is a single-body operator on spin $p$, and the
coefficients $c_{pq}$ satisfy
\begin{align}
  c_p = \sum_q c_{pq},
  &&
  c_{pq} = c_{qp},
  &&
  c_{pp} = 0.
\end{align}
The effective Hamiltonian $H_W$ induced on the ground-state manifold
$\M_0$ by a perturbation $\Z_W$ through second order in perturbation
theory is given by\cite{bravyi2011schrieffer, perlin2019effective}
\begin{align}
  H_W = H_W^{(1)} + H_W^{(2)},
  &&
  H_W^{(1)} = \P_0 \Z_W \P_0,
  &&
  H_W^{(2)} = - \P_0 \Z_W \E \Z_W \P_0,
  &&
  \E \equiv \sum_{k>0} \f{\P_k}{\Delta_k},
\end{align}
where $\P_k$ is a projector onto the $k$-th excited eigenspace of the
interaction Hamiltonian $H_{\t{int}}$, with interaction energy
$\Delta_k$ above that of fully symmetric manifold $\M_0$.  In order to
calculate $H_W$, we will make heavy use of
\begin{enumerate*}
\item the permutational symmetry of the ground-state manifold, and
\item the fact that sandwiching any operator $\O$ between the
  projectors $\P_0$ and $\P_k$ (i.e.~with $k>0$) annihilates any part
  of $\O$ that is not strictly off-diagonal in the subspaces preserved
  by $\P_0$ and $\P_k$.
\end{enumerate*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Single-body perturbation}

We first compute the effective Hamiltonian induced by a sum over
single-body terms, $\Z_1$.  We therefore expand
\begin{align}
  H_1^{(1)} = \sum_p c_p \P_0 Z_p \P_0,
\end{align}
where the permutational symmetry of the ground-state manifold implies
that
\begin{align}
  H_1^{(1)} =  \sum_p c_p \P_0 Z_0 \P_0 = c \P_0 Z \P_0,
  &&
  c \equiv \f1N \sum_p c_p,
  &&
  Z = \sum_p Z_p.
\end{align}
In order to compute $\H_1^{(1)}$, we first expand, for an arbitrary
fully symmetric state $\ket\psi\in\M_0$
\begin{align}
  \H \Z_1 \ket\psi
  = \sum_{\substack{k\\p<q}} c_k \Pi_{pq} Z_k \ket\psi
  = \sum_{\substack{p<q\\k\notin\set{p,q}}} c_k \Pi_{pq} Z_k \ket\psi
  + \sum_{\substack{p<q\\k\in\set{p,q}}} c_k \Pi_{pq} Z_k \ket\psi.
\end{align}
If $k\notin\set{p,q}$, then the permutation operator $\Pi_{pq}$
commutes with $Z_k$ and annihilates on the fully symmetric state
$\ket\psi$, so
\begin{align}
  \sum_{\substack{p<q\\k\notin\set{p,q}}}
  c_k \Pi_{pq} Z_k \ket\psi
  = \sum_{\substack{k\\p<q\\p,q\ne k}}
  c_k Z_k \ket\psi
  = {N-1 \choose 2} \Z_1 \ket\psi.
\end{align}
In the case of $k\in\set{p,q}$, meanwhile,
\begin{align}
  \sum_{\substack{p<q\\k\in\set{p,q}}} c_k \Pi_{pq} Z_k \ket\psi
  = \sum_{k<q} c_k Z_q \ket\psi
  + \sum_{p<k} c_k Z_p \ket\psi
  = \sum_{k\ne q} c_k Z_q \ket\psi,
\end{align}
where
\begin{align}
  \sum_{k\ne q} c_k = \sum_k c_k - c_q = N c - c_q,
\end{align}
so
\begin{align}
  \sum_{\substack{p<q\\k\in\set{p,q}}} c_k \Pi_{pq} Z_k \ket\psi
  = - \sum_q c_q Z_q \ket\psi + N c \sum_q Z_q \ket\psi
  = - \Z_1 \ket\psi + c N Z \ket\psi.
\end{align}
Altogether, we have found that
\begin{align}
  \H \Z_1 \ket\psi
  =  E_1 \Z_1 \ket\psi + c N Z \ket\psi,
  &&
  E_1 \equiv {N-1 \choose 2} - 1
\end{align}
where $E_1$ is an eigenvalue of $\H$ corresponding to the manifold
$\M_1$ of spin-wave states.  We now make the ansatz
\begin{align}
  \Z_1 \ket\psi = \ket\phi = \phi_1 \ket{1} + \phi_0 \ket{0},
  &&
  \ket{0} = c N Z \ket\psi,
\end{align}
where $\ket{k}\in\M_k$ are eigenvectors of $\H$ with eigenvalues
$E_k$.  We then have that
\begin{align}
  \H \ket\phi
  = E_1 \ket\phi + \ket{0}
  = E_1 \phi_1 \ket{1} + \p{E_1 \phi_0 + 1} \ket{0}
  = E_1 \phi_1 \ket{1} + E_0 \phi_0 \ket{0},
\end{align}
which implies that $\phi_0=\p{E_0-E_1}^{-1}$, and in turn
\begin{align}
  \phi_1 \ket{1}
  = \ket\phi - \phi_0 \ket{0}
  = \Z_1 \ket\psi - \f{c N Z}{E_0-E_1} \ket\psi.
\end{align}
Using the fact that $E_0-E_1=N$, we thus discover that the vector
$\tilde\Z_1 \ket\psi$, with
\begin{align}
  \tilde\Z_1 \equiv \Z_1 - c Z = \sum_k \p{c_k-c} Z_k,
\end{align}
is an eigenvector of $\H$ with eigenvalue $E_1$.  Equivalently,
$\tilde\Z_1\ket\psi$ is an eigenvector of the interaction Hamiltonian
$H_{\t{int}}$ with energy $Nu$ above the ground-state energy, which
implies that
\begin{align}
  \E \Z_1 \P_0 = \E \tilde\Z_1 \P_0 = \f1{Nu} \tilde\Z_1 \P_0,
\end{align}
where we may freely take $\Z_1\to\tilde\Z_1$ above because the
difference $\Z_1-\tilde\Z_1$ is a permutationally symmetric operator
that is annihilated between the projectors in $\E$ and $\P_0$.  The
second-order effective Hamiltonian induced by a single-body
perturbation is thus
\begin{align}
  H_1^{(2)} = -\P_0 \Z_1 \E \Z_1 \P_0
  = -\P_0 \tilde\Z_1 \E \tilde\Z_1 \P_0
  = - \f1{Nu} \P_0 \tilde\Z_1 \tilde\Z_1 \P_0.
\end{align}
We now need only to simplify the product
\begin{align}
  \P_0 \tilde\Z_1 \tilde\Z_1 \P_0
  = \sum_{p,q} \p{c_p-c} \p{c_q-c} \P_0 Z_p Z_q \P_0.
\end{align}
This product is straightforward to simplify by using the permutational
symmetry of the ground-state manifold.  As an example, we consider
single-body operators $Z_p$ square to the identity, in which case the $p=q$ terms above are
\begin{align}
  \sum_p \p{c_p-c}^2 \P_0 Z_p^2 \P_0
  = N \var_p c_p \P_0,
  &&
  \var_p c_p \equiv \f1N \sum_p \p{c_p-c}^2.
\end{align}
Similarly, the $p\ne q$ terms are
\begin{align}
  \sum_{p\ne q} \p{c_p-c} \p{c_q-c} \P_0 Z_p Z_q \P_0
  \sum_p \p{c_p-c} \sum_{q\ne p} \p{c_q-c} \P_0 Z_0 Z_1 \P_0,
\end{align}
where
\begin{align}
  \sum_{q\ne p} \p{c_q-c} = \sum_q \p{c_q-c} - \p{c_p-c} = - \p{c_p-c},
\end{align}
so
\begin{align}
  \sum_p \p{c_p-c} \sum_{q\ne p} \p{c_q-c} = - N \var_p c_p,
\end{align}
and
\begin{align}
  \P_0 Z Z \P_0
  = \sum_p \P_0 Z_p^2 \P_0
  + \sum_{p\ne q} \P_0 Z_p Z_q \P_0
  = N \P_0 + N \p{N-1} \P_0 Z_0 Z_1 \P_0,
\end{align}
so
\begin{align}
  \sum_{p\ne q} \p{c_p-c} \p{c_q-c} \P_0 Z_p Z_q \P_0
  = - \f{\var_p c_p}{N-1}\p{\P_0 Z Z \P_0 - N \P_0}.
\end{align}
Altogether, if the single-body operators $Z_p$ square to the identity, then
\begin{align}
  \P_0 \tilde\Z_1 \tilde\Z_1 \P_0
  = - \f{\var_p c_p}{N-1} \p{\P_0 Z Z \P_0 - N^2 \P_0},
\end{align}
and in turn
\begin{align}
  H_1^{(2)} = \f{u}{N\p{N-1}} \var_p c_p \p{\P_0 Z Z \P_0 - N^2 \P_0}
  \simeq \f{u}{N\p{N-1}} \var_p c_p ZZ,
\end{align}
where $\simeq$ denotes equality up to
\begin{enumerate*}
\item a restrict to the ground-state manifold $\M_0$, and
\item overall constants that have no physical consequence.
\end{enumerate*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Two-body perturbation}

The first-order effective Hamiltonian $H_2^{(1)}$ induced by the
two-body perturbation $\Z_2$ is straightforward to calculate by making
use of the permutational symmetry of the ground-state manifold.  We
therefore jump straight into the calculation of the second-order
effective Hamiltonian $H_2^{(2)}$.  As before, we pick an arbitrary
fully symmetric state $\ket\psi\in\M_0$ and expand
\begin{align}
  \H \Z_2 \ket\psi
  = \f12 \sum_{\substack{k\ne\ell\\p<q}}
  c_{k\ell} \Pi_{pq} Z_k Z_\ell \ket\psi,
\end{align}
where the permutation operator has no effect in terms with
$p,q\notin\set{k,\ell}$ or $\set{p,q}=\set{k,\ell}$, so
\begin{align}
  \H \Z_2 \ket\psi
  = \sp{{N-2 \choose 2} + 1} \Z_2 \ket\psi
  + \f12 \sum_{\substack{k,\ell\\k\ne q\ne\ell}}
  c_{k\ell} Z_q Z_\ell \ket\psi
  + \f12 \sum_{\substack{k,\ell\\\ell\ne q\ne k}}
  c_{k\ell} Z_k Z_q \ket\psi.
\end{align}
We can then simplify
\begin{align}
  \sum_{k\ne q} c_{k\ell}
  = \sum_k c_{k\ell} - c_{q\ell}
  = c_\ell - c_{q\ell},
\end{align}
which implies that
\begin{align}
  \sum_{\substack{k,\ell\\k=p\ne q\ne\ell}}
  c_{k\ell} Z_q Z_\ell \ket\psi
  = - \sum_{q\ne\ell} c_{q\ell} Z_q Z_\ell
  + \sum_{q\ne\ell} c_\ell Z_q Z_\ell
  = - \Z_2 + Z \Z_1 - \sum_\ell c_\ell Z_\ell^2,
\end{align}
and similarly
\begin{align}
  \sum_{\substack{k,\ell\\\ell\ne q\ne k}} c_{k\ell} Z_k Z_q \ket\psi
  = - \Z_2 + \Z_1 Z - \sum_k c_k Z_k^2.
\end{align}
We thus have that
\begin{align}
  \H \Z_2 \ket\psi = \sp{{ N-2\choose 2 } - 1} \Z_2 \ket\psi
  + \f12 \sp{Z,\Z_1}_+ \ket\psi - \sum_\ell c_\ell Z_\ell^2,
\end{align}
where $\sp{X,Y}_+ \equiv X Y + Y X$ is an anti-commutator.  Similarly
to the single-body case, we now make the ansatz
\begin{align}
  \Z_2 \ket\psi = \ket\phi
  = \phi_2 \ket{2} + \phi_1 \ket{1} + \phi_0 \ket{0},
\end{align}
where
\begin{align}
  \ket{1} = \f12 \sp{Z,\Z_1}_+ \ket\psi,
  &&
  \ket{0} = - \sum_\ell c_\ell Z_\ell^2 \ket\psi,
\end{align}
and $\ket{k}\in\M_k$ is an eigenvector of the $k$-th excited manifold
of the interaction Hamiltonian $H_{\t{int}}=-u\H$, with corresponding
eigenvalue $E_k$ with respect to $\H$.  We then have that
\begin{align}
  \H \ket\phi
  = E_2 \phi_2 \ket{2}
  + \p{E_2\phi_1 + 1} \ket{1}
  + \p{E_2\phi_0 + 1}\ket{0}
  = E_2 \phi_2 \ket{2} + E_1 \phi_1 \ket{2} + E_0 \phi_0 \ket{0},
\end{align}
which implies
\begin{align}
  \phi_1 = \f1{E_1-E_2} = \f1{N-2},
  &&
  \phi_0 = \f1{E_0-E_2} = \f1{2\p{N-1}},
\end{align}
and in turn
\begin{align}
  \phi_2 \ket{2}
  = \ket\phi - \phi_1 \ket{1} - \phi_0 \ket{0}.
\end{align}
We thus find that the vector $\tilde\Z_2\ket\psi$, with
\begin{align}
  \tilde\Z_2 \equiv \Z_2 - \f1{N-2} \times \f12 \sp{Z,\Z_1}_+
  + \f1{2\p{N-1}} \sum_\ell c_\ell Z_\ell^2,
\end{align}
is an eigenvector of $\H$ with eigenvalue $E_2={N-2 \choose 2}-1$.  If
the single-body operators square to the identity, then
\begin{align}
  \tilde\Z_2
  \equiv \Z_2 - \f1{N-2} \times \f12 \sp{Z,\Z_1}_+
  + \f{c}{2\p{N-1}},
  &&
  c = \f1{N^2} \sum_{p,q} c_{pq}.
\end{align}

\bibliography{multilevel_spin_notes.bib}

\end{document}

### 2019-10-28 collective_sun/multilevel_spin_notes.tex
cleanup / corrections in appendix on raman drive coefficients

@@
 algebra provided in Appendix \ref{sec:trans_prod}.  For clarity, if
-any coefficient $\Omega_{LM}^{(\phi,j)}$ is independent of $\phi$, we
-suppress the explicit dependence on $\phi$, i.e.~taking
-$\Omega_{LM}^{(\phi,j)}\to\Omega_{LM}$.  Defining the re-scaled
-driving amplitudes
-\begin{align}
+any coefficient $\Omega_{LM}^{(\phi,j)}$ is independent of the SOC
+angle $\phi$ or lattice site $j$, we suppress the explicit dependence
+on $\p{\phi,j}$, i.e.~taking $\Omega_{LM}^{(\phi,j)}\to\Omega_{LM}$.
+Defining the re-scaled driving amplitudes
+\begin{align}
+  \tilde\Omega_{LM}^{(\phi,j)} \equiv \xi_L^{-1} \Omega_{LM}^{(\phi,j)},
+  &&
   \tilde\Omega_m \equiv \xi_1^{-1} \Omega_m,
-  &&
-  \tilde\Omega_{LM} \equiv \xi_L^{-1} \Omega_{LM},
-\end{align}
-where the scale factors $\xi_L$ are defined in \eqref{eq:scale_fac},
-the coefficients of the effective Raman drive in
-\eqref{eq:raman_drive_apndx} are determined by
+\end{align}
+where where the scale factors $\xi_L$ are defined in
+\eqref{eq:scale_fac} and $\Omega_m$ are the (real) driving amplitudes
+in the ``bare'' Raman drive in \eqref{eq:raman_bare}, the coefficients
+of the effective Raman drive in \eqref{eq:raman_drive_apndx} are
+determined by
 \begin{align}
@@
 \end{align}
-Here $\tilde\Omega_m$, $\phi_m$, and $\Delta$ are respectively the
-(real) driving amplitudes, phases, and detuning that appear in the
-``bare'' Raman drive in \eqref{eq:raman_bare}.  If the driving lasers
-are phase-matched with $\phi_m=m\phi$, such that a nuclear spin
-transition $\mu\to\mu+m$ on site $j$ imprints the phase $m\phi j$,
-then the drive $H_{\t{Raman}}^{(\phi)}$ takes the form
+Here $\phi_m$ and $\Delta$ are respectively the phases and detuning
+that appear in the ``bare'' Raman drive in \eqref{eq:raman_bare}.  If
+the driving lasers are phase-matched with $\phi_m=m\phi$, such that a
+nuclear spin transition $\mu\to\mu+m$ on site $j$ imprints the phase
+$m\phi j$, then the drive $H_{\t{Raman}}^{(\phi)}$ takes the form
 \begin{align}

### 2019-10-28 soc_squeezing/squeezing_paper/main.tex
final changes to squeezing paper prior to publication

@@
 Optical lattice clocks in particular have seen some of the most impressive developments in recent years, reaching record levels of precision ($\sim 3\times 10^{-19}$)\cite{campbell2017fermidegenerate, marti2018imaging} and accuracy ($\sim 1\times 10^{-18}$)\cite{bloom2014optical, mcgrew2018atomic}.
-These advancements required important breakthroughs, including the capability to cool and trap fermionic alkaline-earth atoms in spin-insensitive potentials\cite{takamoto2003spectroscopy, barber2006direct, ye2008quantum}; the development of ultra-coherent lasers\cite{kessler2012sub40mhzlinewidth, cole2013tenfold, matei2017mu} to fully exploit an ultra-narrow clock transition\cite{ludlow2015optical}; the detailed characterization of inter-atomic interactions\cite{scazza2014observation, cappellini2014direct, zhang2014spectroscopic}; and, more recently, the preparation of a quantum degenerate gas in a three-dimensional (3D) optical lattice\cite{campbell2017fermidegenerate, marti2018imaging, goban2018emergence}.
+These advancements required important breakthroughs, including the capability to cool and trap fermionic alkaline-earth atoms in spin-insensitive potentials\cite{takamoto2003spectroscopy, barber2006direct, ye2008quantum}; the development of ultracoherent lasers\cite{kessler2012sub40mhzlinewidth, cole2013tenfold, matei2017mu} to fully exploit an ultranarrow clock transition\cite{ludlow2015optical}; the detailed characterization of inter-atomic interactions\cite{scazza2014observation, cappellini2014direct, zhang2014spectroscopic}; and, more recently, the preparation of a quantum degenerate gas in a three-dimensional (3D) optical lattice\cite{campbell2017fermidegenerate, marti2018imaging, goban2018emergence}.
 Nonetheless, all improvements in sensing capabilities to date have been based on single-particle control of internal atomic degrees of freedom.
@@
 The major challenge for progress in this direction is that generating entanglement requires interactions, which are generally undesirable because they degrade atomic coherence, thereby limiting clock performance\cite{swallows2011suppression, martin2013quantum, rey2014probing, ludlow2011coldcollisionshift, lemke2011wave, ludlow2015optical}.
-In fact, the most precise and accurate optical lattice clocks were designed to operate with fermionic atoms in identical nuclear and electronic states in order to suppress collisional decoherence\cite{campbell2009probing, swallows2011suppression, campbell2017fermidegenerate}, as identical fermions cannot interact via the otherwise dominant ($s$-wave) collisions at ultracold temperatures.
+In fact, the most precise and accurate optical lattice clocks were designed to operate with fermionic atoms in identical nuclear and electronic states to suppress collisional decoherence\cite{campbell2009probing, swallows2011suppression, campbell2017fermidegenerate}, as identical fermions cannot interact via the otherwise dominant ($s$-wave) collisions at ultracold temperatures.
 However, an initially spin-polarized Fermi gas still exhibits interactions at later times due to spin-orbit coupling (SOC) that is induced by the laser that drives the clock transition (i.e.~the ``clock laser'')\cite{wall2016synthetic, kolkowitz2016spinorbitcoupled, livi2016synthetic, bromley2018dynamics}.
@@
 Additionally, we show that by applying a modulated drive from the clock laser, one can further prepare states that saturate the Heisenberg limit $\Delta\theta\sim1/N$ for phase sensitivity\cite{kitagawa1993squeezed, ma2011quantum, degen2017quantum}.
-This capability mirrors efforts in other settings, such as NV centers\cite{bauch2018ultralong, aiello2013compositepulse} and trapped ions\cite{burd2019quantum}, to enhance quantum metrology through the use of driven non-equilibrium phenomena.
+This capability mirrors efforts in other settings, such as nitrogen-vacancy centers in diamond\cite{bauch2018ultralong, aiello2013compositepulse} and trapped ions\cite{burd2019quantum}, to enhance quantum metrology through the use of driven non-equilibrium phenomena.
 
@@
 where $\overline B\equiv\sum_q B_q/N$ is the mean and $\widetilde B^2\equiv\sum_q\p{B_q-\overline B}^2/N$ the variance of the axial field.
-The effect of the $\sim\v S\c\v S$ term is to generate a relative phase between states with different total spin $S$ and thus has no effect on dynamics  restricted to a fixed $S$.
+The effect of the $\sim\hat{\v S}\c\hat{\v S}$ term is to generate a relative phase between states with different total spin $S$ and thus has no effect on dynamics  restricted to a fixed $S$.
 Note also that  the collective spin rotation from $\overline B \hat S_\z$  can be eliminated by going into a rotating frame or by using a spin echo.
@@
 Further improvements upon this scheme can be made by introducing a time-dependent driving field that transforms the OAT Hamiltonian into a two-axis twisting (TAT) one.
-While the OAT model initially generates squeezing faster than the TAT model, the squeezing generation rate of OAT (measured in dB per second) falls off with time, while the squeezing generation rate for TAT remains approximately constant until reaching Heisenberg-limited amount of spin squeezing with $\xi^2_{\t{opt}}\sim N^{-1}$ (see Appendix \ref{sec:time_series})\cite{kitagawa1993squeezed}.
+While the OAT model initially generates squeezing faster than the TAT model, the squeezing generation rate of OAT (measured in dB per second) falls off with time, while the squeezing generation rate for TAT remains approximately constant until reaching Heisenberg-limited amount of spin squeezing with $\xi^2_{\t{opt}}\sim N^{-1}$\cite{kitagawa1993squeezed}.
 
@@
 The strength of SOC for Raman transitions is tunable and, moreover, the lifetimes of ground nuclear spin levels are longer than 100 seconds in the lattice\cite{goban2018emergence}.
-In this case, our protocol for preparing a squeezed state would additionally end with a coherent state transfer from nuclear to electronic degrees of freedom in order to retain metrological utility for the atomic clock.
+In this case, our protocol for preparing a squeezed state would additionally end with a coherent state transfer from nuclear to electronic degrees of freedom to retain metrological utility for the atomic clock.
 If, for example, the $-9/2$ and $-7/2$ nuclear spin states are used for the preparation of a squeezed state, then the collective-spin entanglement of atoms can be transferred to electronic states at the end of the spin squeezing protocol with a $\sigma^-$ polarized $\pi$ pulse.
@@
 We acknowledge helpful discussions with M.~Norcia, C.~Sanner, and M.~Mamaev.
-This work is supported by the Air Force Office of Scientific Research (AFOSR) grant FA9550-18-1-0319; the AFOSR Multidisciplinary University Research Initiative (MURI) grant; the Defense Advanced Research Projects Agency (DARPA) and Army Research Office (ARO) grant W911NF-16-1-0576; the National Science Foundation (NSF) grant PHY-1820885; JILA-NSF grant PFC-173400; and the
-National Institute of Standards and Technology (NIST).
+This work is supported by the Air Force Office of Scientific Research (AFOSR) grant FA9550-18-1-0319; the AFOSR Multidisciplinary University Research Initiative (MURI) grant; the Defense Advanced Research Projects Agency (DARPA) and Army Research Office (ARO) grant W911NF-16-1-0576; the National Science Foundation (NSF) grant PHY-1820885; JILA-NSF grant PFC-1734006; and the National Institute of Standards and Technology (NIST).
 
@@
   = -\f1{fU} \sum_{m,k,n,n'} b_n b_{n'}
-  \bk{m | s_\z^{(n)} |mk} \\
-  \times \bk{mk | s_\z^{(n')} | m},
+  \op{m} s_\z^{(n)} \ket{mk} \\
+  \times \bra{mk} s_\z^{(n')} \op{m},
   \label{eq:general_H_eff_2}
@@
 For $N = 60$, we have filled all delocalized modes as well as several localized modes, resulting in a large region of no density fluctuations at the trap center.
-Panel (d) contains the eigenspectrum for a single internal state in the presence of the trap (with the index $n$ labelling the eigenvalues in order of increasing energy), where the critical mode $n_c$ dividing the spatially delocalized and localized modes is indicated  by a black dash-dotted line.
+Panel (d) contains the eigenspectrum for a single internal state in the presence of the trap (with the index $n$ labeling the eigenvalues in order of increasing energy), where the critical mode $n_c$ dividing the spatially delocalized and localized modes is indicated  by a black dash-dotted line.
 The highest occupied mode in the $\downarrow$ ground state for $N = 20$ and $N = 60$ is indicated by the green and red solid lines, respectively.}

### 2019-10-30 collective_sun/sun_perturbation.tex
changing notes on perturbation theory with SU(n)-symmetric interactions to the case of non-uniform couplings

\documentclass[nofootinbib,notitlepage,11pt]{revtex4-2}

%%% linking references
\usepackage{hyperref}
\hypersetup{
  breaklinks=true,
  colorlinks=true,
  linkcolor=blue,
  filecolor=magenta,
  urlcolor=cyan,
}

%%% header / footer
\usepackage{fancyhdr} % easier header and footer management
\pagestyle{fancy} % page formatting style
\fancyhf{} % clear all header and footer text
\renewcommand{\headrulewidth}{0pt} % remove horizontal line in header
\usepackage{lastpage} % for referencing last page
\cfoot{\thepage~of \pageref{LastPage}} % "x of y" page labeling

%%% symbols, notations, etc.
\usepackage{physics,braket,bm,amssymb} % physics and math
\renewcommand{\t}{\text} % text in math mode
\newcommand{\f}[2]{\dfrac{#1}{#2}} % shorthand for fractions
\newcommand{\p}[1]{\left(#1\right)} % parenthesis
\renewcommand{\sp}[1]{\left[#1\right]} % square parenthesis
\renewcommand{\set}[1]{\left\{#1\right\}} % curly parenthesis
\newcommand{\bk}{\Braket} % shorthand for braket notation
\renewcommand{\v}{\bm} % bold vectors
\newcommand{\uv}[1]{\bm{\hat{#1}}} % unit vectors
\newcommand{\av}{\vec} % arrow vectors
\renewcommand{\d}{\text{d}} % for infinitesimals
\renewcommand{\c}{\cdot} % inner product

\usepackage{dsfont} % for identity operator
\newcommand{\1}{\mathds{1}}

\newcommand{\up}{\uparrow}
\newcommand{\dn}{\downarrow}

\newcommand{\x}{\text{x}}
\newcommand{\y}{\text{y}}
\newcommand{\z}{\text{z}}

\newcommand{\B}{\mathcal{B}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathcal{E}}
\renewcommand{\H}{\mathcal{H}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\M}{\mathcal{M}}
\newcommand{\N}{\mathcal{N}}
\renewcommand{\O}{\mathcal{O}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\T}{\mathcal{T}}
\renewcommand{\S}{\mathcal{S}}
\newcommand{\V}{\mathcal{V}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\Z}{\mathcal{Z}}

\newcommand{\EE}{\mathbb{E}}
\renewcommand{\SS}{\mathbb{S}}
\newcommand{\ZZ}{\mathbb{Z}}

\newcommand{\FS}{\text{FS}}

\DeclareMathOperator{\sign}{sign}
\DeclareMathOperator{\cov}{cov}
\let\var\relax
\DeclareMathOperator{\var}{var}

\def\obra#1{\mathinner{({#1}|}}
\def\oket#1{\mathinner{|{#1})}}
\def\obk#1{\mathinner{({#1})}}
\def\oop#1#2{\oket{#1}\!\obra{#2}}

\usepackage[inline]{enumitem} % in-line lists and \setlist{} (below)
\setlist[enumerate,1]{label={(\roman*)}} % default in-line numbering
\setlist{nolistsep} % more compact spacing between environments

%%% text markup
\usepackage{color} % text color
\newcommand{\red}[1]{{\color{red} #1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\thispagestyle{fancy}

\title{Perturbing SU($n$)-symmetric interactions}%
\author{Michael A. Perlin}%
\date{\today}

\maketitle

We consider an array of $N$ multilevel spins with non-uniform
SU($n$)-symmetric interactions that can be written in the form
\begin{align}
  H_0 = \sum_{p<q} h_{pq} \Pi_{pq},
  &&
  \Pi_{pq} \equiv \sum_{\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)},
\end{align}
where $S_{\mu\nu}^{(p)}\equiv\op{\mu}{\nu}_p$ flips the state of spin
$p$ to $\ket\mu$ from $\ket\nu$; the operator $\Pi_{pq}$ permutes
spins $p$ and $q$, and the coefficients $h_{pq}$ satisfy
\begin{align}
  h_{pp} = 0
  &&
  h_{pq} \le 0,
  &&
  h_{pq} = h_{qp}.
\end{align}
The ground-state manifold of the interaction Hamiltonian $H_0$
consists of fully symmetric states that are simultaneous $+1$
eigenstates of all permutation operators $\Pi_{pq}$.  The energy of
these states is
\begin{align}
  E_0 \equiv \sum_{p<q} h_{pq} = \f12 \sum_{p,q} h_{pq}
  = \f12 \sum_p h_p = \f12 N h,
\end{align}
where we define
\begin{align}
  h_p \equiv \sum_q h_{pq},
  &&
  h \equiv \EE_p\sp{h_p} = \f1N \sum_{p,q} h_{pq},
  &&
  \EE_p\sp{X_p} \equiv \f1N \sum_p X_p.
\end{align}
We wish to determine the effective dynamics induced on the manifold
$\M_0$ of fully symmetric states by weak perturbations of the form
\begin{align}
  \V_1 \equiv \sum_p v_p V_p,
  &&
  \V_2 \equiv \sum_{p<q} v_{pq} V_p V_q,
  \label{eq:perturbations}
\end{align}
where $V_p$ is a single-body operator on spin $p$, and
$v_{pq}=v_{qp}$.  The effective Hamiltonian $H_W$ induced on the
ground-state manifold $\M_0$ by a perturbation $\V_W$ through second
order in perturbation theory is given by\cite{bravyi2011schrieffer,
  perlin2019effective}
\begin{align}
  H_W = H_W^{(1)} + H_W^{(2)},
  &&
  H_W^{(1)} = \P_0 \V_W \P_0,
  &&
  H_W^{(2)} = - \P_0 \V_W \E \V_W \P_0,
  &&
  \E \equiv \sum_{k>0} \f{\P_k}{\Delta_k},
\end{align}
where $\P_k$ is a projector onto the $k$-th excited eigenspace of the
interaction Hamiltonian $H_0$, with interaction energy $\Delta_k$
above that of fully symmetric manifold $\M_0$.

As an immediate disclaimer, we note that these notes do {\bf not}
solve the general problem of calculating $H_1,H_2$ for arbitrary
perturbations $\V_1,\V_2$.  Instead, we determine conditions under
which the perturbations $\V_1,\V_2$ generate eigenvectors of the
interaction Hamiltonian $H_0$, and calculate $H_1$ in the case that
these conditions are satisfied.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Single-body perturbations}

We first compute the effective Hamiltonian induced by the single-body
perturbation $\V_1$.  We therefore expand
\begin{align}
  H_1^{(1)} = \sum_p v_p \P_0 V_p \P_0,
\end{align}
and use the permutational symmetry of the ground-state manifold to
simplify
\begin{align}
  H_1^{(1)} =  \sum_p v_p \P_0 V_0 \P_0 \simeq v V,
  &&
  v \equiv \EE_p\sp{v_p},
  &&
  V = \sum_p V_p,
  \label{eq:H_1_1}
\end{align}
where $\simeq$ denotes equality up to a restriction to the
ground-state manifold $\M_0$.  In order to compute the second order
effective Hamiltonian $\H_1^{(2)}$, we first choose an arbitrary fully
symmetric state $\ket\psi\in\M_0$ and expand
\begin{align}
  H_0 \V_1 \ket\psi
  = \sum_{\substack{k\\p<q}} h_{pq} v_k \Pi_{pq} V_k \ket\psi
  = \sum_{\substack{p<q\\k\notin\set{p,q}}}
  h_{pq} v_k \Pi_{pq} V_k \ket\psi
  + \sum_{\substack{p<q\\k\in\set{p,q}}}
  h_{pq} v_k \Pi_{pq} V_k \ket\psi.
  \label{eq:H_V1_psi_start}
\end{align}
In the case of $k\notin\set{p,q}$, the permutation operator $\Pi_{pq}$
commutes with $V_k$ and annihilates on $\ket\psi$, leaving us the sum
\begin{align}
  \sum_{\substack{p<q\\k\notin\set{p,q}}} h_{pq}
  = \f12 \sum_{\substack{p,q\\k\notin\set{p,q}}} h_{pq}
  = \f12 \sum_{p,q} h_{pq}
  - \f12 \sum_{\substack{p,q\\k\in\set{p,q}}} h_{pq}
  = \f12 N h - \sum_q h_{kq}
  = E_0 - h_k,
\end{align}
where we have used the facts that $h_{pp}=0$ and $h_{pq}=h_{qp}$.  The
terms with $k\notin\set{p,q}$ in \eqref{eq:H_V1_psi_start} are then
\begin{align}
  \sum_{\substack{p<q\\k\notin\set{p,q}}}
  h_{pq} v_k \Pi_{pq} V_k \ket\psi
  = \sum_k \p{E_0 - h_k} v_k V_k \ket\psi
  = E_0 \V_1 \ket\psi - \sum_k h_k v_k V_k \ket\psi.
\end{align}
In the case of $k\in\set{p,q}$, we have
\begin{align}
  \sum_{\substack{p<q\\k\in\set{p,q}}}
  h_{pq} v_k \Pi_{pq} V_k \ket\psi
  = \sum_{k<q} h_{kq} v_k V_q \ket\psi
  + \sum_{p<k} h_{pk} v_k V_p \ket\psi
  = \sum_{k,q} h_{kq} v_k V_q \ket\psi,
\end{align}
so in total,
\begin{align}
  H_0 \V_1 \ket\psi
  = E_0 \V_1 \ket\psi + \sum_{q,k} h_{kq} \p{v_k - v_q} V_q \ket\psi.
  \label{eq:H_V1_psi}
\end{align}
Denoting a vector of the coefficients $v_q$ by $\v v$, a matrix of all
coupling constants $h_{pq}$ by $\v h$, and a diagonal matrix with
$h_q\equiv\sum_k h_{kq}$ on the diagonals by $\tilde{\v h}$, we thus
find that the vector $\V_1\ket\psi$ is an eigenvector of the
interaction Hamiltonian $H_0$ if the coefficients $\v v$ satisfy the
eigenvalue equation
\begin{align}
  \p{\v h - \tilde{\v h}}\c\v v = \Delta\p{\v v} \v v
  \label{eq:v_eig_1}
\end{align}
where $\v x\c\v y$ denotes a standard matrix product of $\v x$ and
$\v y$, and $\Delta\p{\v v}$ is an eigenvalue of
$\v h - \tilde{\v h}$.  If the interaction Hamiltonian $H_0$ is
translationally invariant, then all $h_q=h$, and $\tilde{\v h}=h\,\1$
is proportional to the identity matrix $\1$, so the condition in
\eqref{eq:v_eig_1} becomes
\begin{align}
  \v h \c\v v = \sp{\Delta\p{\v v}-h} \v v.
\end{align}
If we construct a perturbation $\V_1\p{\v v}$ with coefficients $\v v$
that satisfy \eqref{eq:v_eig_1}, then
\begin{align}
  H_0 \V_1\p{\v v} \ket\psi
  = \sp{E_0 + \Delta\p{\v v}} \V_1\p{\v v} \ket\psi,
\end{align}
for any fully symmetric state $\ket\psi\in\M_0$, which for
$\Delta\p{\v v}>0$ implies that
\begin{align}
  \E \V_1\p{\v v} \P_0
  = \sum_{k>0} \f1{\Delta_k} \P_k \V_1\p{\v v} \P_0
  = \f1{\Delta\p{\v v}} \V_1\p{\v v} \P_0,
\end{align}
and in turn
\begin{align}
  H_1^{(2)}
  = -\P_0 \V_1\p{\v v} \E \V_1\p{\v v} \P_0
  = -\f1{\Delta\p{\v v}} \P_0 \V_1\p{\v v}^2 \P_0.
\end{align}
Computing the second-order effective Hamiltonian $H_1^{(2)}$ thus
requires us to simplify the product
\begin{align}
  \P_0 \V_1\p{\v v}^2 \P_0
  = \sum_{p,q} v_p v_q \P_0 V_p V_q \P_0
  = \sum_p v_p^2 \P_0 V_p^2 \P_0
  + \sum_{p\ne q} v_p v_q \P_0 V_p V_q \P_0.
  \label{eq:PVVP_start}
\end{align}
We note that vectors $\v v$ satisfying \eqref{eq:v_eig_1} with
$\Delta\p{\v v}\ne 0$ must be mean-zero, as
\begin{align}
  v = \f1N \sum_q v_q
  = \f{\sum_{q,k} h_{qk} v_k}{N\sp{\Delta\p{\v v}+h}}
  = \f{\sum_k h_k v_k}{N\sp{\Delta\p{\v v}+h}}
  = \f{h v}{\Delta\p{\v v}+h}
  = \f{v}{\Delta\p{\v v}/h+1},
\end{align}
which is only possible if $\Delta\p{\v v}=0$.  It follows that the
first sum on the right of \eqref{eq:PVVP_start} is
\begin{align}
  \sum_p v_p^2 \P_0 V_p^2 \P_0
  = \sum_p v_p^2 \P_0 V_0^2 \P_0
  = \var\p{\v v} \P_0 W \P_0,
  &&
  W \equiv \sum_p V_p^2,
\end{align}
where
\begin{align}
  \var\p{\v x} \equiv \EE_p\sp{\p{x_p-\EE_q\sp{x_q}}^2}
  = \EE_p\sp{x_p^2} - \EE_q\sp{x_q}^2
\end{align}
is the variance of the elements in the vector $\v x$.  Similarly, the
second sum on the right of \eqref{eq:PVVP_start} is
\begin{align}
  \sum_{p\ne q} v_p v_q \P_0 V_p V_q \P_0
  = \sum_{p\ne q} v_p v_q \P_0 V_0 V_1 \P_0,
  \label{eq:PVVP_neq_start}
\end{align}
where if the coefficients in $\v v$ are mean-zero, then
\begin{align}
  \sum_{p\ne q} v_p v_q
  = \sum_p v_p \sum_{q\ne p} v_q
  = \sum_p v_p \p{\sum_q v_q - v_p}
  = - \sum_P v_p^2
  = - N \var\p{\v v}.
\end{align}
To evaluate the operator content of \eqref{eq:PVVP_neq}, we expand
\begin{align}
  \P_0 V^2 \P_0
  = \sum_p \P_0 V_p^2 \P_0
  + \sum_{p\ne q} \P_0 V_p V_q \P_0
  = \P_0 W \P_0 + N \p{N-1} \P_0 V_0 V_1 \P_0,
\end{align}
which implies that
\begin{align}
  \sum_{p\ne q} v_p v_q \P_0 V_p V_q \P_0
  = -\f{\var\p{\v v}}{N-1}\P_0\p{V^2 - W} \P_0.
  \label{eq:PVVP_neq}
\end{align}
Altogether, we have that
\begin{align}
  \P_0 \V_1\p{\v v}^2 \P_0
  = -\f{\var\p{\v v}}{N-1} \P_0 \p{V^2 - N W} \P_0,
\end{align}
and so
\begin{align}
  H_1^{(2)} = \f{\var\p{\v v}}{\p{N-1}\Delta\p{\v v}} \p{V^2 - N W},
\end{align}
where we implicitly restrict the effective Hamiltonian $H_1^{(2)}$ to
the ground-state manifold $\M_0$, and note that the result above holds
only for
\begin{enumerate*}
\item translationally invariant Hamiltonians, and
\item perturbations $\V_1\p{\v v}$ constructed from vectors $\v v$
  that satisfy the eigenvalue equation in \eqref{eq:v_eig_1}.
\end{enumerate*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Two-body perturbations}

The first-order effective Hamiltonian $H_2^{(1)}$ induced on the fully
symmetric manifold $\M_0$ by the two-body perturbation in $\V_2$ in
\eqref{eq:perturbations} is
\begin{align}
  H_2^{(1)}
  = \P_0 \V_2 \P_0
  = \sum_{p<q} v_{pq} \P_0 V_p V_q \P_0
  = \f12 \sum_{p\ne q} v_{pq} \P_0 V_p V_q \P_0
  \simeq -\f12 \f{\var\p{\v v}}{N-1} \p{V^2 - W},
\end{align}
where we made use of the result in \eqref{eq:PVVP_neq}.  In order to
compute the second-order effective Hamiltonian $H_2^{(2)}$, as before
we pick an arbitrary fully symmetric state $\ket\psi\in\M_0$ and
expand
\begin{align}
  H \V_2 \ket\psi
  = \sum_{\substack{k<\ell\\p<q}} v_{k\ell} h_{pq}
  \Pi_{pq} V_k V_\ell \ket\psi.
  \label{eq:H_V2_psi_start}
\end{align}
The permutation operator $\Pi_{pq}$ acts trivially on
$V_k V_\ell \ket\psi$ when $p,q\notin\set{k,\ell}$ or
$\set{p,q}=\set{k,\ell}$, leaving a sum of the form
\begin{align}
  \sum_{\substack{p<q\\p,q\notin\set{k,\ell}}} h_{pq} + h_{k\ell}
  = \sum_{p<q} h_{pq}
  - \sum_{\substack{p<q\\p\in\set{k,\ell}\\q\notin\set{k,\ell}}} h_{pq}
  - \sum_{\substack{p<q\\p\notin\set{k,\ell}\\q\in\set{k,\ell}}} h_{pq}
  = E_0 - \sum_{\substack{p\in\set{k,\ell}\\q\notin\set{k,\ell}}} h_{pq}
  = E_0 - 2 h + 2 h_{k\ell},
\end{align}
where we assumed translational invariance with all $h_p=h$.  The
corresponding terms in \eqref{eq:H_V2_psi_start} are then
\begin{align}
  \sum_{k<\ell}
  \sp{\sum_{\substack{p<q\\p,q\notin\set{k,\ell}}} h_{pq} + h_{k\ell}}
  v_{k\ell} V_k V_\ell \ket\psi
  = \p{E_0-2h} \V_2 \ket\psi
  + \sum_{k,\ell} h_{k\ell} v_{k\ell} V_k V_\ell \ket\psi.
\end{align}
The terms in \eqref{eq:H_V2_psi_start} with only one of
$p,q\in\set{k,\ell}$, meanwhile, reduce to
\begin{align}
  \f12 \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}}
  v_{k\ell} \p{h_{kq} V_q V_\ell + h_{\ell q} V_k V_q} \ket\psi
  = \sum_{k,\ell,q} v_{k\ell} h_{kq} V_q V_\ell \ket\psi
  - \sum_{k,\ell} v_{k\ell} h_{k\ell} V_k^2 \ket\psi.
\end{align}
In total,
\begin{align}
  H_0 \V_2 \ket\psi
  = \p{E_0 - 2h} \V_2 \ket\psi
  + \sum_{k,\ell} \p{h_{k\ell} v_{k\ell} + \sum_q h_{kq} v_{q\ell}}
  V_k V_\ell \ket\psi
  - \sum_{k,\ell} h_{k\ell} v_{k\ell} V_k^2 \ket\psi.
  \label{eq:H_V2_psi_mid}
\end{align}
We thus find that the vector $\V_2\ket\psi$ is an eigenvector of the
interaction Hamiltonian $H_0$ if the matrix $\v v$ of all $v_{k\ell}$
satisfies, for some constant $\varepsilon\p{\v v}$,
\begin{align}
  \v h\circ \v v + \v h\c\v v
  = \varepsilon\p{\v v} \v v,
  \label{eq:v_eig_2}
\end{align}
where $\v h\circ\v v$ denotes an element-wise (Hadamard) product of
the matrices $\v h$ and $\v v$, i.e.~with components
$h_{k\ell} v_{k\ell}$; and $\v h\c\v v$ denotes an ordinary matrix
product, with components $\sum_q h_{kq} v_{q\ell}$.  Note that the
diagonal components of the condition in \eqref{eq:v_eig_2} enforce
that the last sum in \eqref{eq:H_V2_psi_mid} vanishes.  If we
construct a perturbation $\V_2\p{\v v}$ with coefficients $\v v$ that
satisfy \eqref{eq:v_eig_2}, then
\begin{align}
  H_0 \V_2\p{\v v} \ket\psi
  = \sp{E_0 - 2h + 2\varepsilon\p{\v v}} \V_2\p{\v v} \ket\psi.
\end{align}

\bibliography{multilevel_spin_notes.bib}

\end{document}

### 2019-10-30 collective_sun/sun_perturbation.tex
more discussion / derivations in sun_perturbation.tex

@@
 \newcommand{\bk}{\Braket} % shorthand for braket notation
-\renewcommand{\v}{\bm} % bold vectors
-\newcommand{\uv}[1]{\bm{\hat{#1}}} % unit vectors
-\newcommand{\av}{\vec} % arrow vectors
-\renewcommand{\d}{\text{d}} % for infinitesimals
+
 \renewcommand{\c}{\cdot} % inner product
+\renewcommand{\oc}{\circ} % element-wise product
+
+\newcommand{\m}{\bm} % bold symbol
+\renewcommand{\v}{\vec} % arrow vector
 
@@
 
+\renewcommand{\d}{\text{d}} % for infinitesimals
 \newcommand{\x}{\text{x}}
@@
 \DeclareMathOperator{\var}{var}
+\DeclareMathOperator{\diag}{diag}
 
@@
 As an immediate disclaimer, we note that these notes do {\bf not}
-solve the general problem of calculating $H_1,H_2$ for arbitrary
-perturbations $\V_1,\V_2$.  Instead, we determine conditions under
-which the perturbations $\V_1,\V_2$ generate eigenvectors of the
-interaction Hamiltonian $H_0$, and calculate $H_1$ in the case that
-these conditions are satisfied.
+solve the general problem of calculating effective Hamiltonians
+$H_1,H_2$ for arbitrary perturbations $\V_1,\V_2$.  Instead, we
+\begin{enumerate*}
+\item calculate the first-order effective Hamiltonians
+  $H_1^{(1)},H_2^{(1)}$,
+\item determine conditions under which the perturbations $\V_1,\V_2$
+  generate eigenvectors of the interaction Hamiltonian $H_0$, and
+\item calculate $H_1$ in the case that these conditions are satisfied
+  by $\V_1$.
+\end{enumerate*}
+When the conditions for $\V_1,\V_2$ to generate eigenvectors of $H_0$
+are not satisfied, we hope that the techniques in these notes can be
+used to construct eigenvectors and eigenvalues of $H_0$ that will
+enable the use of ordinary perturbation theory to determine the
+effective Hamiltonians $H_1,H_2$ induced on the fully symmetric
+manifold $\M_0$ by the perturbations $\V_1,\V_2$.
 
@@
 Denoting a vector of the coefficients $v_q$ by $\v v$, a matrix of all
-coupling constants $h_{pq}$ by $\v h$, and a diagonal matrix with
-$h_q\equiv\sum_k h_{kq}$ on the diagonals by $\tilde{\v h}$, we thus
-find that the vector $\V_1\ket\psi$ is an eigenvector of the
-interaction Hamiltonian $H_0$ if the coefficients $\v v$ satisfy the
-eigenvalue equation
+coupling constants $h_{pq}$ by $\m h$, and a vector of all
+$h_q\equiv\sum_k h_{kq}$ by $\v h$, we thus find that the vector
+$\V_1\ket\psi$ is an eigenvector of the interaction Hamiltonian $H_0$
+if the coefficients $\v v$ satisfy
 \begin{align}
-  \p{\v h - \tilde{\v h}}\c\v v = \Delta\p{\v v} \v v
-  \label{eq:v_eig_1}
+  \m h \c \v v - \v h_\d\oc\v v = \Delta\p{\v v} \v v,
+  \label{eq:v_cond_1}
 \end{align}
-where $\v x\c\v y$ denotes a standard matrix product of $\v x$ and
-$\v y$, and $\Delta\p{\v v}$ is an eigenvalue of
-$\v h - \tilde{\v h}$.  If the interaction Hamiltonian $H_0$ is
-translationally invariant, then all $h_q=h$, and $\tilde{\v h}=h\,\1$
-is proportional to the identity matrix $\1$, so the condition in
-\eqref{eq:v_eig_1} becomes
+where $\m h\c\v v$ is an ordinary matrix-vector product of $\m h$ and
+$\v v$, with components $\sum_q h_{kq} v_q$; $\v h_\d\oc\v v$ is an
+element-wise (Hadamard) product of $\v h_\d$ and $\v v$, with
+components $h_k v_k$; and $\Delta\p{\v v}$ is some constant that may
+depend on $\v v$.  Finding vectors $\v v$ that satisfy
+\eqref{eq:v_cond_1} is equivalent to finding eigenvectors of the
+matrix $\m h-\m h_\d$, where $\m h_\d\equiv\diag\v h_\d$ is a diagonal
+matrix; the constant $\Delta\p{\v v}$ is then the eigenvalue of
+$\m h-\m h_\d$ corresponding to the eigenvector $\v v$.  If the
+interaction Hamiltonian $H_0$ is translationally invariant, then all
+$h_q=h$ and $\v h_\d\oc\v v=h\v v$, so the condition in
+\eqref{eq:v_cond_1} simplifies to
 \begin{align}
-  \v h \c\v v = \sp{\Delta\p{\v v}-h} \v v.
+  \m h \c\v v = \sp{\Delta\p{\v v}+h} \v v.
 \end{align}
 If we construct a perturbation $\V_1\p{\v v}$ with coefficients $\v v$
-that satisfy \eqref{eq:v_eig_1}, then
+that satisfy \eqref{eq:v_cond_1}, then
 \begin{align}
@@
 \end{align}
-We note that vectors $\v v$ satisfying \eqref{eq:v_eig_1} with
-$\Delta\p{\v v}\ne 0$ must be mean-zero, as
+We note that vectors $\v v$ satisfying \eqref{eq:v_cond_1} and
+$\Delta\p{\v v}\ne 0$ must be mean-zero, as evident by summing over
+the components of \eqref{eq:v_cond_1}:
 \begin{align}
-  v = \f1N \sum_q v_q
-  = \f{\sum_{q,k} h_{qk} v_k}{N\sp{\Delta\p{\v v}+h}}
-  = \f{\sum_k h_k v_k}{N\sp{\Delta\p{\v v}+h}}
-  = \f{h v}{\Delta\p{\v v}+h}
-  = \f{v}{\Delta\p{\v v}/h+1},
+  \Delta\p{\v v} \sum_k v_k
+  = \sum_{k,q} h_{kq} v_q - \sum_k h_k v_k
+  = \sum_q h_q v_q - \sum_k h_k v_k
+  = 0.
+  \label{eq:D_v_0}
 \end{align}
-which is only possible if $\Delta\p{\v v}=0$.  It follows that the
-first sum on the right of \eqref{eq:PVVP_start} is
+It follows that the first sum on the right of \eqref{eq:PVVP_start} is
 \begin{align}
@@
   = \sum_p v_p \p{\sum_q v_q - v_p}
-  = - \sum_P v_p^2
+  = - \sum_p v_p^2
   = - N \var\p{\v v}.
 \end{align}
-To evaluate the operator content of \eqref{eq:PVVP_neq}, we expand
+To evaluate the operator content of \eqref{eq:PVVP_neq_start}, we
+expand
 \begin{align}
@@
   = \P_0 W \P_0 + N \p{N-1} \P_0 V_0 V_1 \P_0,
+  \label{eq:PVVP_neq_ops}
 \end{align}
@@
 the ground-state manifold $\M_0$, and note that the result above holds
-only for
-\begin{enumerate*}
-\item translationally invariant Hamiltonians, and
-\item perturbations $\V_1\p{\v v}$ constructed from vectors $\v v$
-  that satisfy the eigenvalue equation in \eqref{eq:v_eig_1}.
-\end{enumerate*}
+only for perturbations $\V_1\p{\v v}$ constructed from vectors $\v v$
+that satisfy the eigenvalue equation in \eqref{eq:v_cond_1}.
 
@@
   = \sum_{p<q} v_{pq} \P_0 V_p V_q \P_0
-  = \f12 \sum_{p\ne q} v_{pq} \P_0 V_p V_q \P_0
-  \simeq -\f12 \f{\var\p{\v v}}{N-1} \p{V^2 - W},
+  \simeq \f{v}{N-1} \p{V^2 - W},
+  &&
+  v \equiv \f1N \sum_{p<q} v_{pq},
 \end{align}
-where we made use of the result in \eqref{eq:PVVP_neq}.  In order to
-compute the second-order effective Hamiltonian $H_2^{(2)}$, as before
-we pick an arbitrary fully symmetric state $\ket\psi\in\M_0$ and
-expand
+where we made use of the previous result in \eqref{eq:PVVP_neq_ops}.
+In order to compute the second-order effective Hamiltonian
+$H_2^{(2)}$, as before we pick an arbitrary fully symmetric state
+$\ket\psi\in\M_0$ and expand
 \begin{align}
@@
   = \sum_{\substack{k<\ell\\p<q}} v_{k\ell} h_{pq}
+  \Pi_{pq} V_k V_\ell \ket\psi
+  = \f12 \sum_{\substack{k,\ell\\p<q}} v_{k\ell} h_{pq}
   \Pi_{pq} V_k V_\ell \ket\psi.
@@
 $V_k V_\ell \ket\psi$ when $p,q\notin\set{k,\ell}$ or
-$\set{p,q}=\set{k,\ell}$, leaving a sum of the form
+$\set{p,q}=\set{k,\ell}$, leaving a sum over $p,q$ of the form
 \begin{align}
@@
   = E_0 - \sum_{\substack{p\in\set{k,\ell}\\q\notin\set{k,\ell}}} h_{pq}
-  = E_0 - 2 h + 2 h_{k\ell},
+  = E_0 + 2 h_{k\ell} - h_k - h_\ell.
 \end{align}
-where we assumed translational invariance with all $h_p=h$.  The
-corresponding terms in \eqref{eq:H_V2_psi_start} are then
+The corresponding terms in \eqref{eq:H_V2_psi_start} with
+$p,q\notin\set{k,\ell}$ and $\set{p,q}=\set{k,\ell}$ are then
 \begin{align}
-  \sum_{k<\ell}
-  \sp{\sum_{\substack{p<q\\p,q\notin\set{k,\ell}}} h_{pq} + h_{k\ell}}
-  v_{k\ell} V_k V_\ell \ket\psi
-  = \p{E_0-2h} \V_2 \ket\psi
-  + \sum_{k,\ell} h_{k\ell} v_{k\ell} V_k V_\ell \ket\psi.
+  E_0 \V_2 \ket\psi + \f12 \sum_{k,\ell}
+  \p{2 h_{k\ell} - h_k - h_\ell} v_{k\ell} V_k V_\ell \ket\psi,
 \end{align}
-The terms in \eqref{eq:H_V2_psi_start} with only one of
-$p,q\in\set{k,\ell}$, meanwhile, reduce to
+while the terms in \eqref{eq:H_V2_psi_start} with only one of
+$p,q\in\set{k,\ell}$ simplify to
 \begin{align}
@@
   H_0 \V_2 \ket\psi
-  = \p{E_0 - 2h} \V_2 \ket\psi
-  + \sum_{k,\ell} \p{h_{k\ell} v_{k\ell} + \sum_q h_{kq} v_{q\ell}}
-  V_k V_\ell \ket\psi
+  = E_0 \V_2 \ket\psi
+  + \sum_{k,\ell} \p{\sp{h_{k\ell}-\p{\f{h_k+h_\ell}{2}}} v_{k\ell}
+    + \sum_q h_{kq} v_{q\ell}} V_k V_\ell \ket\psi
   - \sum_{k,\ell} h_{k\ell} v_{k\ell} V_k^2 \ket\psi.
@@
 We thus find that the vector $\V_2\ket\psi$ is an eigenvector of the
-interaction Hamiltonian $H_0$ if the matrix $\v v$ of all $v_{k\ell}$
-satisfies, for some constant $\varepsilon\p{\v v}$,
+interaction Hamiltonian $H_0$ if the matrix $\m v$ of all $v_{k\ell}$
+satisfies, for some constant $\varepsilon\p{\m v}$ that may depend on
+$\m v$,
 \begin{align}
-  \v h\circ \v v + \v h\c\v v
-  = \varepsilon\p{\v v} \v v,
-  \label{eq:v_eig_2}
+  \tilde{\m h}\oc\m v + \m h\c\m v
+  = \varepsilon\p{\m v} \m v,
+  &&
+  \tilde{\m h}
+  \equiv \m h - \f12\p{\v h_\d\otimes\v\1 + \v\1\otimes\v h_\d},
+  \label{eq:v_cond_2}
 \end{align}
-where $\v h\circ\v v$ denotes an element-wise (Hadamard) product of
-the matrices $\v h$ and $\v v$, i.e.~with components
-$h_{k\ell} v_{k\ell}$; and $\v h\c\v v$ denotes an ordinary matrix
-product, with components $\sum_q h_{kq} v_{q\ell}$.  Note that the
-diagonal components of the condition in \eqref{eq:v_eig_2} enforce
-that the last sum in \eqref{eq:H_V2_psi_mid} vanishes.  If we
-construct a perturbation $\V_2\p{\v v}$ with coefficients $\v v$ that
-satisfy \eqref{eq:v_eig_2}, then
+where $\tilde{\m h}\oc\m v$ denotes an element-wise (Hadamard) product
+of $\tilde{\m h}$ and $\m v$, with components
+$\tilde h_{k\ell} v_{k\ell}$; $\m h\c\m v$ denotes an ordinary matrix
+product of $\m h$ and $\m v$, with components
+$\sum_q h_{kq} v_{q\ell}$; $\v\1$ is a constant vector of ones; and
+$\v h_\d\otimes\v\1$ is a tensor product of $\v h_\d$ and $\v\1$.
+Note that the diagonal components of the condition in
+\eqref{eq:v_cond_2} automatically enforce that the last sum in
+\eqref{eq:H_V2_psi_mid} vanishes.  The condition in
+\eqref{eq:v_cond_2} can be recast as an eigenvalue equation by
+treating the $N\times N$ matrix $\m v$ as a vector with $N^2$
+components, which can be further reduced to ${N\choose2}=N\p{N-1}/2$
+components by the fact that the matrix $\m v$ is symmetric with zeros
+on the diagonal.  If the interaction Hamiltonian $H_0$ is
+translationally invariant, then all $h_q=h$ and
+$\v h_\d\otimes\v\1\oc\m v=\v\1\otimes\v h_\d\oc\m v=h\m v$, so the
+condition in \eqref{eq:v_cond_2} becomes
 \begin{align}
-  H_0 \V_2\p{\v v} \ket\psi
-  = \sp{E_0 - 2h + 2\varepsilon\p{\v v}} \V_2\p{\v v} \ket\psi.
+  \m h\oc\m v + \m h\c\m v = \sp{\varepsilon\p{\m v}+h} \m v,
 \end{align}
+If we construct a perturbation $\V_2\p{\m v}$ with coefficients $\m v$
+that satisfy \eqref{eq:v_cond_2}, then
+\begin{align}
+  H_0 \V_2\p{\m v} \ket\psi
+  = \sp{E_0 + 2\varepsilon\p{\m v}} \V_2\p{\m v} \ket\psi.
+\end{align}
+As a curiosity, we note that if $\V_1\p{\v v}\ket\psi$ is an
+eigenvector of $H_0$, then the coefficient vector $\v v$ must satisfy
+\eqref{eq:D_v_0}, which we used to determine that $\v v$ was mean-zero
+if $\Delta\p{\v v}\ne0$; the analogous condition on matrices $\m v$ in
+the two-body case is
+\begin{align}
+  N \varepsilon\p{\m v} v = \f12 \sum_{k,\ell} h_{k\ell} v_{k\ell},
+\end{align}
+where $v\equiv\sum_{p<q}v_{pq}/N$ is the mean value of the
+coefficients in $\V_2\p{\m v}$.
 

### 2019-10-30 collective_sun/sun_perturbation.tex
added section on finding eigenvectors of the non-uniform interaction Hamiltonian

@@
 
-\renewcommand{\d}{\text{d}} % for infinitesimals
+\renewcommand{\d}{\text{d}}
 \newcommand{\x}{\text{x}}
@@
 \newcommand{\z}{\text{z}}
+
+\newcommand{\e}{\varepsilon}
 
@@
 components $h_k v_k$; and $\Delta\p{\v v}$ is some constant that may
-depend on $\v v$.  Finding vectors $\v v$ that satisfy
-\eqref{eq:v_cond_1} is equivalent to finding eigenvectors of the
-matrix $\m h-\m h_\d$, where $\m h_\d\equiv\diag\v h_\d$ is a diagonal
-matrix; the constant $\Delta\p{\v v}$ is then the eigenvalue of
-$\m h-\m h_\d$ corresponding to the eigenvector $\v v$.  If the
-interaction Hamiltonian $H_0$ is translationally invariant, then all
-$h_q=h$ and $\v h_\d\oc\v v=h\v v$, so the condition in
-\eqref{eq:v_cond_1} simplifies to
+depend on $\v v$.  If the interaction Hamiltonian $H_0$ is
+translationally invariant, then all $h_q=h$ and
+$\v h_\d\oc\v v=h\v v$, so the condition in \eqref{eq:v_cond_1}
+simplifies to
 \begin{align}
@@
 interaction Hamiltonian $H_0$ if the matrix $\m v$ of all $v_{k\ell}$
-satisfies, for some constant $\varepsilon\p{\m v}$ that may depend on
-$\m v$,
+satisfies, for some constant $\e\p{\m v}$ that may depend on $\m v$,
 \begin{align}
-  \tilde{\m h}\oc\m v + \m h\c\m v
-  = \varepsilon\p{\m v} \m v,
+  \bar{\m h}\oc\m v + \m h\c\m v = \e\p{\m v} \m v,
   &&
-  \tilde{\m h}
+  \bar{\m h}
   \equiv \m h - \f12\p{\v h_\d\otimes\v\1 + \v\1\otimes\v h_\d},
@@
 \end{align}
-where $\tilde{\m h}\oc\m v$ denotes an element-wise (Hadamard) product
-of $\tilde{\m h}$ and $\m v$, with components
+where $\bar{\m h}\oc\m v$ denotes an element-wise (Hadamard) product
+of $\bar{\m h}$ and $\m v$, with components
 $\tilde h_{k\ell} v_{k\ell}$; $\m h\c\m v$ denotes an ordinary matrix
@@
 \eqref{eq:v_cond_2} automatically enforce that the last sum in
-\eqref{eq:H_V2_psi_mid} vanishes.  The condition in
-\eqref{eq:v_cond_2} can be recast as an eigenvalue equation by
-treating the $N\times N$ matrix $\m v$ as a vector with $N^2$
-components, which can be further reduced to ${N\choose2}=N\p{N-1}/2$
-components by the fact that the matrix $\m v$ is symmetric with zeros
-on the diagonal.  If the interaction Hamiltonian $H_0$ is
-translationally invariant, then all $h_q=h$ and
+\eqref{eq:H_V2_psi_mid} vanishes.  If the interaction Hamiltonian
+$H_0$ is translationally invariant, then all $h_q=h$ and
 $\v h_\d\otimes\v\1\oc\m v=\v\1\otimes\v h_\d\oc\m v=h\m v$, so the
@@
 \begin{align}
-  \m h\oc\m v + \m h\c\m v = \sp{\varepsilon\p{\m v}+h} \m v,
+  \m h\oc\m v + \m h\c\m v = \sp{\e\p{\m v}+h} \m v,
 \end{align}
@@
   H_0 \V_2\p{\m v} \ket\psi
-  = \sp{E_0 + 2\varepsilon\p{\m v}} \V_2\p{\m v} \ket\psi.
+  = \sp{E_0 + 2\e\p{\m v}} \V_2\p{\m v} \ket\psi.
 \end{align}
@@
 \begin{align}
-  N \varepsilon\p{\m v} v = \f12 \sum_{k,\ell} h_{k\ell} v_{k\ell},
+  N \e\p{\m v} v = \f12 \sum_{k,\ell} h_{k\ell} v_{k\ell},
 \end{align}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Finding eigenstates of the interaction Hamiltonian}
+
+Here we discuss the construction of eigenstates of the interaction
+Hamiltonian $H_0$.  If a vector $\v v$ and matrix $\m v$ satisfy the
+conditions in \eqref{eq:v_cond_1} and \eqref{eq:v_cond_2}, then the
+single- and two-body operators $\V_1\p{\v v}$ and $\V_2\p{\m v}$
+generate eigenvectors of the interaction Hamiltonian $H_0$ when they
+act on any fully symmetric state $\ket\psi\in\M_0$.  We can use this
+fact to construct eigenstates of $H_0$ by finding solutions to
+\eqref{eq:v_cond_1} and \eqref{eq:v_cond_2}.
+
+The single-body case is straightforward: finding vectors $\v v$ that
+satisfy \eqref{eq:v_cond_1} is equivalent to finding eigenvectors of
+the matrix $\m h-\m h_\d$, where $\m h_\d\equiv\diag\v h_\d$ is a
+diagonal matrix.  The constant $\Delta\p{\v v}$ is then the eigenvalue
+of $\m h-\m h_\d$ that corresponds to the eigenvector $\v v$.  The
+task of constructing eigenvectors $\V_1\p{\v v}\ket\psi$ of $H_0$ and
+determining their energy thus reduces to finding the eigenvectors and
+eigenvalues of the $N\times N$ matrix $\m h-\m h_\d$.
+
+The two-body case requires a bit of work to re-cast as an
+eigenvector/eigenvalue problem.  We begin by writing the matrix $\m v$ as a vector,
+\begin{align}
+  \m v = \sum_{p,q} v_{pq} \op{k}{q}
+  \to \check{\m v} \equiv \sum_{p,q} v_{pq} \ket{pq},
+\end{align}
+and then defining the matrix
+\begin{align}
+  \check{\m h}
+  \equiv \sum_{k,\ell} \bar h_{k\ell} \op{k\ell}
+  + \sum_{k,q,\ell} h_{kq} \op{k}{q} \otimes \op{\ell},
+  \label{eq:h_super_mat}
+\end{align}
+in terms of which the condition in \eqref{eq:v_cond_2} becomes
+\begin{align}
+  \check{\m h} \c \check{\m v} = \e\p{\m v} \check{\m v}.
+\end{align}
+Note that the matrix $\check{\m h}$ in \eqref{eq:h_super_mat} is
+block-diagonal, with $N$ blocks of size $N\times N$.  The task of
+constructing eigenvectors $\V_2\p{\m v}\ket\psi$ of $H_0$ and
+determining their energy thus reduces to finding the eigenvectors and
+eigenvalues of $N$ matrices of size $N\times N$.
+
 \bibliography{multilevel_spin_notes.bib}

### 2019-10-30 collective_sun/sun_perturbation.tex
better commentary on the block-diagonal form of the two-body eigenvalue problem added simplifications of the two-body eigenvalue problem in the case of translational invariance

@@
 The two-body case requires a bit of work to re-cast as an
-eigenvector/eigenvalue problem.  We begin by writing the matrix $\m v$ as a vector,
+eigenvector/eigenvalue problem.  We first write the matrix $\m v$ as a
+vector,
 \begin{align}
   \m v = \sum_{p,q} v_{pq} \op{k}{q}
-  \to \check{\m v} \equiv \sum_{p,q} v_{pq} \ket{pq},
+  \to \v{\m v} \equiv \sum_{p,q} v_{pq} \ket{pq},
 \end{align}
-and then defining the matrix
+and define the matrix
 \begin{align}
@@
   \equiv \sum_{k,\ell} \bar h_{k\ell} \op{k\ell}
-  + \sum_{k,q,\ell} h_{kq} \op{k}{q} \otimes \op{\ell},
+  + \sum_{k,q,\ell} h_{kq} \op{k\ell}{q\ell},
   \label{eq:h_super_mat}
@@
 \begin{align}
-  \check{\m h} \c \check{\m v} = \e\p{\m v} \check{\m v}.
+  \check{\m h} \c \v{\m v} = \e\p{\m v} \v{\m v}.
 \end{align}
-Note that the matrix $\check{\m h}$ in \eqref{eq:h_super_mat} is
-block-diagonal, with $N$ blocks of size $N\times N$.  The task of
-constructing eigenvectors $\V_2\p{\m v}\ket\psi$ of $H_0$ and
-determining their energy thus reduces to finding the eigenvectors and
-eigenvalues of $N$ matrices of size $N\times N$.
+Note that the matrix $\check{\m h}$ in \eqref{eq:h_super_mat} can be
+written in the block-diagonal form
+\begin{align}
+  \check{\m h} = \sum_\ell \m h_\ell \otimes \op{\ell},
+  &&
+  \m h_\ell
+  \equiv \sum_{k,q}\p{\delta_{kq} \bar h_{k\ell} + h_{kq}} \op{k}{q},
+\end{align}
+where each block $\m h_\ell$ is an $N\times N$ matrix.  We can
+diagonalize these blocks independently, and classify their
+eigenvectors by corresponding eigenvalues.  ``Total'' eigenvectors
+$\v{\m v}$ of $\check{\m h}$ are then essentially concatenations of
+eigenvectors from within each block $\m h_\ell$ that all share the
+same eigenvalue.  The task of constructing eigenvectors
+$\V_2\p{\m v}\ket\psi$ of $H_0$ and determining their energy thus
+reduces to finding the eigenvectors and eigenvalues of $N$ matrices
+$\m h_\ell$ of size $N\times N$.  If the interaction Hamiltonian $H_0$
+is translationally invariant, then
+\begin{align}
+  \m h_{\ell} = \m\omega_\ell \c \m h_0 \c \m\omega_\ell^\dag,
+  &&
+  \m\omega_\ell \equiv \sum_k \op{k+\ell}{k},
+\end{align}
+which implies that we need only find the eigenvectors and eigenvalues
+of a single block $\m h_0$, after which we can determine the
+eigenvectors of any block simply by appropriate applications of the
+shift matrix $\m\omega_\ell$.
 

### 2019-10-30 collective_sun/sun_perturbation.tex
added comment about translational invariance in dimensions greater than 1

@@
 $\m h_\ell$ of size $N\times N$.  If the interaction Hamiltonian $H_0$
-is translationally invariant, then
+is translationally invariant, then $\bar h_{k\ell}=h_{k\ell}-h$, and
+$h_{k\ell}$ depends only on the displacement between the spins indexed
+by $k$ and $\ell$.  The blocks $\m h_\ell$ then take the form
 \begin{align}
@@
   \m\omega_\ell \equiv \sum_k \op{k+\ell}{k},
+  \label{eq:block_shift}
 \end{align}
@@
 eigenvectors of any block simply by appropriate applications of the
-shift matrix $\m\omega_\ell$.
+shift matrix $\m\omega_\ell$.  Note that in dimensions $D>1$, the
+relation in \eqref{eq:block_shift} holds for vectors $k,\ell$ that
+index lattice sites, i.e.~$k=\p{k_1,k_2,\cdots,k_D}$ and likewise with
+$\ell$.
 

### 2019-10-30 collective_sun/sun_perturbation.tex
adding table of contents

@@
 \begin{document}
-\thispagestyle{fancy}
 
@@
 \maketitle
+
+\tableofcontents
+
+\section{Introduction}
 
@@
 $p$ to $\ket\mu$ from $\ket\nu$; the operator $\Pi_{pq}$ permutes
-spins $p$ and $q$, and the coefficients $h_{pq}$ satisfy
+spins $p$ and $q$; and the coefficients $h_{pq}$ satisfy
 \begin{align}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Finding eigenstates of the interaction Hamiltonian}
+\section{Constructing eigenstates of SU($n$)-symmetric interactions}
 
@@
 \begin{align}
-  \m v = \sum_{p,q} v_{pq} \op{k}{q}
-  \to \v{\m v} \equiv \sum_{p,q} v_{pq} \ket{pq},
+  \m v = \sum_{k,\ell} v_{k\ell} \op{k}{\ell}
+  \to \v{\m v} \equiv \sum_{k,\ell} v_{k\ell} \ket{k\ell},
 \end{align}
@@
 diagonalize these blocks independently, and classify their
-eigenvectors by corresponding eigenvalues.  ``Total'' eigenvectors
-$\v{\m v}$ of $\check{\m h}$ are then essentially concatenations of
-eigenvectors from within each block $\m h_\ell$ that all share the
-same eigenvalue.  The task of constructing eigenvectors
+eigenvectors by the corresponding eigenvalues.  ``Total'' eigenvectors
+$\v{\m v}$ of $\check{\m h}$ are then essentially built from
+eigenvectors within each block $\m h_\ell$ that share the same
+eigenvalue.  The task of constructing eigenvectors
 $\V_2\p{\m v}\ket\psi$ of $H_0$ and determining their energy thus
 reduces to finding the eigenvectors and eigenvalues of $N$ matrices
-$\m h_\ell$ of size $N\times N$.  If the interaction Hamiltonian $H_0$
-is translationally invariant, then $\bar h_{k\ell}=h_{k\ell}-h$, and
-$h_{k\ell}$ depends only on the displacement between the spins indexed
-by $k$ and $\ell$.  The blocks $\m h_\ell$ then take the form
+$\m h_\ell$ of size $N\times N$.
+
+If the interaction Hamiltonian $H_0$ is translationally invariant,
+then $\bar h_{k\ell}=h_{k\ell}-h$, and $h_{k\ell}$ depends only on the
+displacement between the spins indexed by $k$ and $\ell$.  The blocks
+$\m h_\ell$ then take the form
 \begin{align}

### 2019-10-30 collective_sun/sun_perturbation.tex
generalized the two-body perturbation to the case of different operators on the addressed spins

@@
   &&
-  \V_2 \equiv \sum_{p<q} v_{pq} V_p V_q,
+  \V_2 \equiv \f12 \sum_{p,q} v_{pq} V_p \tilde V_q,
   \label{eq:perturbations}
 \end{align}
-where $V_p$ is a single-body operator on spin $p$, and
-$v_{pq}=v_{qp}$.  The effective Hamiltonian $H_W$ induced on the
-ground-state manifold $\M_0$ by a perturbation $\V_W$ through second
-order in perturbation theory is given by\cite{bravyi2011schrieffer,
-  perlin2019effective}
+where $V_p$ is a single-body operator on spin $p$, and we enforce
+$v_{pq}=v_{qp}$ with $v_{pp}=0$.  The effective Hamiltonian $H_B$
+induced on the ground-state manifold $\M_0$ by a perturbation $\V_B$
+through second order in perturbation theory is given
+by\cite{bravyi2011schrieffer, perlin2019effective}
 \begin{align}
-  H_W = H_W^{(1)} + H_W^{(2)},
+  H_B = H_B^{(1)} + H_B^{(2)},
   &&
-  H_W^{(1)} = \P_0 \V_W \P_0,
+  H_B^{(1)} = \P_0 \V_B \P_0,
   &&
-  H_W^{(2)} = - \P_0 \V_W \E \V_W \P_0,
+  H_B^{(2)} = - \P_0 \V_B \E \V_B \P_0,
   &&
@@
 symmetric manifold $\M_0$ by the two-body perturbation in $\V_2$ in
-\eqref{eq:perturbations} is
+\eqref{eq:perturbations} can be computed using the same tricks as for
+$H_1^{(2)}$:
 \begin{align}
@@
   = \P_0 \V_2 \P_0
-  = \sum_{p<q} v_{pq} \P_0 V_p V_q \P_0
-  \simeq \f{v}{N-1} \p{V^2 - W},
+  = \f12 \sum_{p,q} v_{pq} \P_0 V_p \tilde V_q \P_0
+  \simeq \f{v}{N-1} \p{\f12 \sp{V,\tilde V}_+ - Z},
+\end{align}
+where $V,\tilde V,Z$ are respectively collective operators that sum
+over the index of $V_p,\tilde V_p,Z_p$, and
+\begin{align}
+  v \equiv \f1N \sum_{p<q} v_{pq},
   &&
-  v \equiv \f1N \sum_{p<q} v_{pq},
+  Z_p \equiv \f12 \sp{V_p, \tilde V_p}_+,
 \end{align}
-where we made use of the previous result in \eqref{eq:PVVP_neq_ops}.
-In order to compute the second-order effective Hamiltonian
-$H_2^{(2)}$, as before we pick an arbitrary fully symmetric state
-$\ket\psi\in\M_0$ and expand
+with $\sp{X,Y}_+\equiv XY+YX$ an anti-commutator.  In order to compute
+the second-order effective Hamiltonian $H_2^{(2)}$, as before we pick
+an arbitrary fully symmetric state $\ket\psi\in\M_0$ and expand
 \begin{align}
   H \V_2 \ket\psi
-  = \sum_{\substack{k<\ell\\p<q}} v_{k\ell} h_{pq}
-  \Pi_{pq} V_k V_\ell \ket\psi
   = \f12 \sum_{\substack{k,\ell\\p<q}} v_{k\ell} h_{pq}
-  \Pi_{pq} V_k V_\ell \ket\psi.
+  \Pi_{pq} V_k \tilde V_\ell \ket\psi.
   \label{eq:H_V2_psi_start}
@@
   E_0 \V_2 \ket\psi + \f12 \sum_{k,\ell}
-  \p{2 h_{k\ell} - h_k - h_\ell} v_{k\ell} V_k V_\ell \ket\psi,
+  \p{2 h_{k\ell} - h_k - h_\ell}
+  v_{k\ell} V_k \tilde V_\ell \ket\psi,
 \end{align}
@@
 \begin{align}
-  \f12 \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}}
-  v_{k\ell} \p{h_{kq} V_q V_\ell + h_{\ell q} V_k V_q} \ket\psi
-  = \sum_{k,\ell,q} v_{k\ell} h_{kq} V_q V_\ell \ket\psi
-  - \sum_{k,\ell} v_{k\ell} h_{k\ell} V_k^2 \ket\psi.
+  \f12 \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}} v_{k\ell}
+  \p{h_{kq} V_q \tilde V_\ell + h_{\ell q} V_k \tilde V_q} \ket\psi
+  = \sum_{k,\ell,q} v_{k\ell} h_{kq} V_q \tilde V_\ell \ket\psi
+  - \sum_{k,\ell} v_{k\ell} h_{k\ell} Z_k \ket\psi.
 \end{align}
@@
   + \sum_{k,\ell} \p{\sp{h_{k\ell}-\p{\f{h_k+h_\ell}{2}}} v_{k\ell}
-    + \sum_q h_{kq} v_{q\ell}} V_k V_\ell \ket\psi
-  - \sum_{k,\ell} h_{k\ell} v_{k\ell} V_k^2 \ket\psi.
+    + \sum_q h_{kq} v_{q\ell}} V_k \tilde V_\ell \ket\psi
+  - \sum_{k,\ell} h_{k\ell} v_{k\ell} V_k \tilde V_k \ket\psi.
   \label{eq:H_V2_psi_mid}

### 2019-10-30 collective_sun/sun_perturbation.tex
fixed minor errors

@@
 of $\bar{\m h}$ and $\m v$, with components
-$\tilde h_{k\ell} v_{k\ell}$; $\m h\c\m v$ denotes an ordinary matrix
+$\bar h_{k\ell} v_{k\ell}$; $\m h\c\m v$ denotes an ordinary matrix
 product of $\m h$ and $\m v$, with components
@@
 \end{align}
-where $v\equiv\sum_{p<q}v_{pq}/N$ is the mean value of the
+where $v\equiv\sum_{p<q}v_{pq}/N$ is the mean value of the pair-wise
 coefficients in $\V_2\p{\m v}$.
@@
 
-The two-body case requires a bit of work to re-cast as an
-eigenvector/eigenvalue problem.  We first write the matrix $\m v$ as a
+In order to re-cast the two-body condition in \eqref{eq:v_cond_1} as
+an eigenvector/eigenvalue problem, we write the matrix $\m v$ as a
 vector,

### 2019-10-31 collective_sun/sun_perturbation.tex
finish calculation of second-order effect of single-body perturbation almost finish the same for a two-body perturbation

--- 2019-10-30 collective_sun/sun_perturbation.tex
@@
 \newcommand{\EE}{\mathbb{E}}
+\newcommand{\RR}{\mathbb{R}}
 \renewcommand{\SS}{\mathbb{S}}
@@
 \end{align}
-The ground-state manifold of the interaction Hamiltonian $H_0$
+The ground-state manifold $\M_0$ of the interaction Hamiltonian $H_0$
 consists of fully symmetric states that are simultaneous $+1$
-eigenstates of all permutation operators $\Pi_{pq}$.  The energy of
-these states is
+eigenstates of all permutation operators $\Pi_{pq}$.  We will assume
+throughout these notes that the interaction Hamiltonian $H_0$ is
+gapped, such that there is a finite energy difference between $\M_0$
+and its orthogonal complement.  The energy of fully symmetric states
+$\ket\psi\in\M_0$ is
 \begin{align}
@@
 \end{align}
-We wish to determine the effective dynamics induced on the manifold
-$\M_0$ of fully symmetric states by weak perturbations of the form
+We wish to determine the effective dynamics induced on the fully
+symmetric manifold $\M_0$ by weak single- and two-body perturbations
+of the form
 \begin{align}
@@
   &&
-  \V_2 \equiv \sum_{p<q} v_{pq} V_p V_q,
+  \V_2 \equiv \sum_{p<q} w_{pq} X_p Y_q
+  = \f12 \sum_{p,q} w_{pq} X_p Y_q,
   \label{eq:perturbations}
 \end{align}
-where $V_p$ is a single-body operator on spin $p$, and
-$v_{pq}=v_{qp}$.  The effective Hamiltonian $H_W$ induced on the
-ground-state manifold $\M_0$ by a perturbation $\V_W$ through second
-order in perturbation theory is given by\cite{bravyi2011schrieffer,
-  perlin2019effective}
+where $V_p,X_p,Y_p$ are trace-zero operators on spin $p$, and we
+enforce $w_{pq}=w_{qp}$ with $w_{pp}=0$.  The effective Hamiltonian
+$H_M$ induced on the ground-state manifold $\M_0$ by an $M$-body
+perturbation $\V_M$ through second order in perturbation theory is
+given by\cite{bravyi2011schrieffer, perlin2019effective}
 \begin{align}
-  H_W = H_W^{(1)} + H_W^{(2)},
+  H_M = H_M^{(1)} + H_M^{(2)},
   &&
-  H_W^{(1)} = \P_0 \V_W \P_0,
+  H_M^{(1)} = \P_0 \V_M \P_0,
   &&
-  H_W^{(2)} = - \P_0 \V_W \E \V_W \P_0,
+  H_M^{(2)} = - \P_0 \V_M \E \V_M \P_0,
   &&
-  \E \equiv \sum_{k>0} \f{\P_k}{\Delta_k},
+  \E \equiv \sum_{\Delta>0} \f{\P_\Delta}{\Delta},
 \end{align}
-where $\P_k$ is a projector onto the $k$-th excited eigenspace of the
-interaction Hamiltonian $H_0$, with interaction energy $\Delta_k$
+where $\P_\Delta$ is a projector onto the eigenspace of the
+interaction Hamiltonian $H_0$, with interaction energy $\Delta$
 above that of fully symmetric manifold $\M_0$.
-
-As an immediate disclaimer, we note that these notes do {\bf not}
-solve the general problem of calculating effective Hamiltonians
-$H_1,H_2$ for arbitrary perturbations $\V_1,\V_2$.  Instead, we
-\begin{enumerate*}
-\item calculate the first-order effective Hamiltonians
-  $H_1^{(1)},H_2^{(1)}$,
-\item determine conditions under which the perturbations $\V_1,\V_2$
-  generate eigenvectors of the interaction Hamiltonian $H_0$, and
-\item calculate $H_1$ in the case that these conditions are satisfied
-  by $\V_1$.
-\end{enumerate*}
-When the conditions for $\V_1,\V_2$ to generate eigenvectors of $H_0$
-are not satisfied, we hope that the techniques in these notes can be
-used to construct eigenvectors and eigenvalues of $H_0$ that will
-enable the use of ordinary perturbation theory to determine the
-effective Hamiltonians $H_1,H_2$ induced on the fully symmetric
-manifold $\M_0$ by the perturbations $\V_1,\V_2$.
 
@@
 
-We first compute the effective Hamiltonian induced by the single-body
-perturbation $\V_1$.  We therefore expand
+To compute the effective Hamiltonian induced by the single-body
+perturbation $\V_1$, we expand
 \begin{align}
@@
 ground-state manifold $\M_0$.  In order to compute the second order
-effective Hamiltonian $\H_1^{(2)}$, we first choose an arbitrary fully
-symmetric state $\ket\psi\in\M_0$ and expand
+effective Hamiltonian $H_1^{(2)}$, we first choose an arbitrary fully
+symmetric state $\ket\psi\in\M_0$ and expand (see Appendix
+\ref{sec:H_V1_psi}):
 \begin{align}
   H_0 \V_1 \ket\psi
-  = \sum_{\substack{k\\p<q}} h_{pq} v_k \Pi_{pq} V_k \ket\psi
-  = \sum_{\substack{p<q\\k\notin\set{p,q}}}
-  h_{pq} v_k \Pi_{pq} V_k \ket\psi
-  + \sum_{\substack{p<q\\k\in\set{p,q}}}
-  h_{pq} v_k \Pi_{pq} V_k \ket\psi.
+  = E_0 \V_1 \ket\psi
+  + \sum_{k,q} \p{h_{kq} v_q - h_k v_k} V_k \ket\psi.
+  \label{eq:H_V1_psi}
+\end{align}
+Denoting a vector of the coefficients $v_q$ by $\v v$, a matrix of all
+coupling constants $h_{pq}$ by $\m h$, and a vector of all
+$h_q\equiv\sum_k h_{kq}$ by $\v h$, we thus find that the vector
+$\V_1\ket\psi$ is an eigenvector of the interaction Hamiltonian $H_0$
+if the coefficients $\v v$ satisfy
+\begin{align}
+  \m h \c \v v - \v h\oc\v v
+  = \p{\m h-\diag\v h}\c\v v
+  = \Delta\p{\v v} \v v,
+  \label{eq:cond_1}
+\end{align}
+where $\m h\c\v v$ is an ordinary matrix-vector product of $\m h$ and
+$\v v$, with components $\sum_q h_{kq} v_q$; $\v h\oc\v v$ is an
+element-wise (Hadamard) product of $\v h$ and $\v v$, with components
+$h_k v_k$; $\diag\v h$ is a matrix with $\v h$ on the diagonal and
+zeroes everywhere else; and $\Delta\p{\v v}$ is some constant that may
+depend on $\v v$.  If the interaction Hamiltonian $H_0$ is
+translationally invariant, then all $h_q=h$ and $\v h\oc\v v=h\v v$,
+so the condition in \eqref{eq:cond_1} simplifies to
+\begin{align}
+  \m h \c\v v = \sp{\Delta\p{\v v}+h} \v v.
+\end{align}
+If we {\it construct} a perturbation $\V_1\p{\v v_\Delta}$ with
+coefficients $\v v_\Delta$ that satisfy \eqref{eq:cond_1} with
+eigenvalue $\Delta$, then
+\begin{align}
+  H_0 \V_1\p{\v v_\Delta} \ket\psi
+  = \sp{E_0 + \Delta} \V_1\p{\v v_\Delta} \ket\psi.
+\end{align}
+Finding operators $\V_1\p{\v v}$ that generate eigenvectors of the
+interaction Hamiltonian $H_0$ when they act on fully symmetric states
+$\ket\psi$ thus reduces to finding eigenvectors $\v v$ of
+$\m h-\diag\v h$.  Interestingly, the energy of the (unnormalized)
+state $\V_1\p{\v v}\ket\psi$ depends only on the coefficients $\v v$,
+and is entirely independent of the state $\ket\psi$ or choice of
+single-spin trace-zero operator used to build $\V_1$ in
+\eqref{eq:perturbations}.
+
+If the interaction Hamiltonian $H_0$ is gapped, then a vector
+$\V_1\p{\v v_\Delta}\ket\psi$ with $\Delta=0$ must lie with in the
+fully symmetric manifold $\M_0$, which implies that the operator
+$\V_1\p{\v v_\Delta}$ preserves the permutational symmetry of
+$\ket\psi$.  The only way for $\V_1\p{\v v_\Delta}$ to obey
+permutational symmetry is for $\v v_\Delta$ to be a constant vector,
+i.e.~$\v v_\Delta=v_\Delta\v\1$ with $\1$ a constant vector of ones.
+Indeed, a constant vector $\v\1$ satisfies the condition in
+\eqref{eq:cond_1} with eigenvalue $0$.  All other vectors
+$\v v_\Delta$ satisfying \eqref{eq:cond_1} with $\Delta\ne0$ must be
+orthogonal to $\v\1$, and therefore mean-zero.
+
+We now return to the task of computing the second-order effective
+Hamiltonian $H_1^{(2)}$.  Any coefficient vector $\v v$ can be
+expanded into its projections $\v v_\Delta$ onto the eigenspace of
+vectors satisfying \eqref{eq:cond_1} with eigenvalue $\Delta$,
+i.e.~$\v v = \sum_\Delta \v v_\Delta$, which also allows us to expand
+\begin{align}
+  \V_1\p{\v v} = \sum_\Delta \V_1\p{\v v_\Delta},
+\end{align}
+where each operator $\V_1\p{\v v_\Delta}$ generates a state with
+interaction energy $\Delta$ above that of the fully symmetric
+manifold.  We can therefore simplify
+\begin{align}
+  H_1^{(2)}
+  = -\P_0 \V_1\p{\v v} \E \V_1\p{\v v} \P_0
+  = - \sum_{\Delta>0} \f1{\Delta}
+  \P_0 \V_1\p{\v v} \P_\Delta \V_1\p{\v v} \P_0
+  = - \sum_{\Delta>0} \f1{\Delta} \P_0 \V_1\p{\v v_\Delta}^2 \P_0,
+\end{align}
+where the the product $\P_0 \V_1\p{\v v_\Delta}^2 \P_0$ is worked out
+in Appendix \ref{sec:PVVP}, giving us
+\begin{align}
+  H_1^{(2)} \simeq \f1{N-1} \sum_{\Delta>0}
+  \f{\var\p{\v v_\Delta}}{\Delta} \p{V^2 - N W},
+  \label{eq:H_1_2}
+\end{align}
+where $\simeq$ denotes equality up to a restriction to the fully
+symmetric manifold $\M_0$, and
+\begin{align}
+  \var\p{\v x} \equiv \EE_p\sp{\p{x_p-\EE_q\sp{x_q}}^2},
+  &&
+  W \equiv \sum_p V_p^2.
+\end{align}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Two-body perturbations}
+
+The first-order effective Hamiltonian $H_2^{(1)}$ induced on the fully
+symmetric manifold $\M_0$ by the two-body perturbation $\V_2$ in
+\eqref{eq:perturbations} can be computed using the permutational
+symmetry of $\M_0$ (see Appendix \ref{sec:PXYP}):
+\begin{align}
+  H_2^{(1)}
+  = \P_0 \V_2 \P_0
+  = \sum_{p<q} w_{pq} \P_0 X_p Y_q \P_0
+  \simeq \f{w}{N-1} \p{\f12 \sp{X,Y}_+ - Z},
+\end{align}
+where
+\begin{align}
+  w \equiv \f1N \sum_{p<q} w_{pq},
+  &&
+  X \equiv \sum_p X_p,
+  &&
+  Y \equiv \sum_p Y_p,
+  &&
+  Z \equiv \f12 \sum_p \sp{X_p, Y_p}_+,
+\end{align}
+and $\sp{X,Y}_+\equiv XY+YX$ is an anti-commutator.  In order to
+compute the second-order effective Hamiltonian $H_2^{(2)}$, as before
+we pick an arbitrary fully symmetric state $\ket\psi\in\M_0$ and
+expand (see Appendix \ref{sec:H_V2_psi})
+\begin{align}
+  H_0 \V_2 \ket\psi
+  = E_0 \V_2 \ket\psi
+  + \sum_{k,\ell} \p{\sp{h_{k\ell}-\p{\f{h_k+h_\ell}{2}}} w_{k\ell}
+    + \sum_q h_{kq} w_{q\ell}} X_k Y_\ell \ket\psi
+  - \sum_{k,\ell} h_{k\ell} w_{k\ell} Z_k \ket\psi,
+  \label{eq:H_V2_psi}
+\end{align}
+where $Z_k\equiv\p{1/2}\sp{X_k,Y_k}_+$.  We thus find that the vector
+$\V_2\ket\psi$ is an eigenvector of the interaction Hamiltonian $H_0$
+if the matrix $\m w$ of all $w_{k\ell}$ satisfies
+\begin{align}
+  \bar{\m h}\oc\m w + \m h\c\m w = \e\p{\m w} \m w,
+  &&
+  \bar{\m h}
+  \equiv \m h - \f12\p{\v h\otimes\v\1 + \v\1\otimes\v h},
+  \label{eq:cond_2}
+\end{align}
+where $\bar{\m h}\oc\m w$ denotes an element-wise (Hadamard) product
+of $\bar{\m h}$ and $\m w$, with components
+$\bar h_{k\ell} w_{k\ell}$; $\m h\c\m w$ denotes an ordinary matrix
+product of $\m h$ and $\m w$, with components
+$\sum_q h_{kq} w_{q\ell}$; $\e\p{\m w}$ is some constant that may
+depend on $\m w$; $\v\1$ is a constant vector of ones; and
+$\v h\otimes\v\1$ is a tensor product of $\v h$ and $\v\1$.  Note that
+the diagonal components of the condition in \eqref{eq:cond_2}
+automatically enforce that the last sum in \eqref{eq:H_V2_psi}
+vanishes.
+
+We can re-cast the condition in \eqref{eq:cond_2} as an eigenvalue
+problem by writing the coefficient matrix $\m w$ as a vector in
+$\RR^{N^2}$,
+\begin{align}
+  \m w = \sum_{k,\ell} w_{k\ell} \op{k}{\ell}
+  \to \v{\m w} \equiv \sum_{k,\ell} w_{k\ell} \ket{k\ell},
+\end{align}
+and defining the matrix
+\begin{align}
+  \check{\m h}
+  \equiv \sum_{k,\ell} \bar h_{k\ell} \op{k\ell}
+  + \sum_{k,q,\ell} h_{kq} \op{k\ell}{q\ell},
+  \label{eq:h_super_mat}
+\end{align}
+in terms of which the condition in \eqref{eq:cond_2} becomes the
+eigenvalue equation
+\begin{align}
+  \check{\m h} \c \v{\m w} = \e\p{\m w} \v{\m w}.
+  \label{eq:cond_2_eig}
+\end{align}
+The matrix $\check{\m h}$ in \eqref{eq:h_super_mat} can be written in
+the block-diagonal form
+\begin{align}
+  \check{\m h} = \sum_\ell \m h_\ell \otimes \op{\ell},
+  &&
+  \m h_\ell
+  \equiv \sum_{k,q}\p{\delta_{kq} \bar h_{k\ell} + h_{kq}} \op{k}{q},
+\end{align}
+which reduces the eigenvalue equation in \eqref{eq:cond_2_eig} to the
+$N$ smaller, uncoupled eigenvalue equations
+\begin{align}
+  \m h_\ell \c \v{\m w}_\ell = \e\p{\m w} \v{\m w}_\ell,
+  &&
+  \v{\m w}_\ell \equiv \sum_k w_{k\ell} \ket{k}.
+  \label{eq:cond_2_eig_block}
+\end{align}
+We can thus (numerically) diagonalize the blocks $\m h_\ell$
+independently, and classify their eigenvectors $\v{\m w}_\ell$ by the
+corresponding eigenvalues $\e\p{\m w}$.  ``Total'' eigenvectors
+$\v{\m w}$ of $\check{\m h}$ are then built from eigenvectors
+$\v{\m w}_\ell$ within each block $\m h_\ell$ that share the same
+eigenvalue $\e\p{\m w}$.  The task of constructing eigenvectors
+$\V_2\p{\m w}\ket\psi$ of $H_0$ and determining their energy thus
+reduces to finding the eigenvectors and eigenvalues of $N$ matrices
+$\m h_\ell$ of size $N\times N$.
+
+If the interaction Hamiltonian $H_0$ is translationally invariant,
+then $\bar h_{k\ell}=h_{k\ell}-h$, and $h_{k\ell}$ depends only on the
+displacement between the spins indexed by $k$ and $\ell$.  The blocks
+$\m h_\ell$ then take the form
+\begin{align}
+  \m h_{\ell} = \m\Sigma_\ell \c \m h_0 \c \m\Sigma_\ell^\dag,
+  &&
+  \m\Sigma_\ell \equiv \sum_k \op{k+\ell}{k},
+  \label{eq:block_shift}
+\end{align}
+which implies that we need only find the eigenvectors and eigenvalues
+of a single block $\m h_0$, after which we can determine the
+eigenvectors of any block simply by appropriate applications of the
+shift matrix $\m\Sigma_\ell$.  Note that in dimensions $D>1$, the
+relation in \eqref{eq:block_shift} holds for vectors $k,\ell$ that
+index lattice sites, i.e.~$k=\p{k_1,k_2,\cdots,k_D}$ and likewise with
+$\ell$.
+
+As with the case of single-body perturbations, we can decompose any
+matrix $\m w$ into its projections $\m w_\e$ onto the eigenspace of
+matrices satisfying \eqref{eq:cond_2} with eigenvalue $\e$,
+i.e.~$\m w=\sum_\e\m w_\e$, which allows us to expand
+\begin{align}
+  \V_2\p{\m w} = \sum_\e \V_2\p{\m w_\e}.
+\end{align}
+We can then simplify
+\begin{align}
+  H_2^{(2)} = - \P_0 \V_2\p{\m w} \E \V_2\p{\m w} \P_0
+  = -\sum_{\e>0} \f1\e \P_0 \V_2\p{\m w_\e} \P_\e \V_2\p{\m w_\e} \P_0
+  = -\sum_{\e>0} \f1\e \P_0 \V_2\p{\m w_\e}^2 \P_0,
+\end{align}
+where the product $\P_0 \V_2\p{\m w_\e}^2 \P_0$ is worked out in
+Appendix \ref{sec:PVVP_2}, giving us \red{[to do: finish calculation]}
+\begin{align}
+  H_2^{(2)} =
+\end{align}
+
+\vspace{3cm}
+
+\appendix
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Diagnosing perturbed states with the interaction Hamiltonian}
+
+Here we simplify vectors of the form $H_0\V_M\ket\psi$ to arrive at
+the expansions in \eqref{eq:H_V1_psi} and \eqref{eq:H_V2_psi}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Single-body perturbation}
+\label{sec:H_V1_psi}
+
+We wish to simplify
+\begin{align}
+  H_0 \V_1 \ket\psi
+  = \sum_{\substack{k\\p<q}} h_{pq} v_k \Pi_{pq} V_k \ket\psi,
   \label{eq:H_V1_psi_start}
 \end{align}
-In the case of $k\notin\set{p,q}$, the permutation operator $\Pi_{pq}$
-commutes with $V_k$ and annihilates on $\ket\psi$, leaving us the sum
+which has terms with $k\in\set{p,q}$, and terms with
+$k\notin\set{p,q}$.  In the case of $k\notin\set{p,q}$, the
+permutation operator $\Pi_{pq}$ commutes with $V_k$ and annihilates on
+$\ket\psi$, leaving us the sum
 \begin{align}
@@
 \end{align}
-so in total,
+which with some re-indexing gives us
 \begin{align}
   H_0 \V_1 \ket\psi
-  = E_0 \V_1 \ket\psi + \sum_{q,k} h_{kq} \p{v_k - v_q} V_q \ket\psi.
-  \label{eq:H_V1_psi}
+  = E_0 \V_1 \ket\psi
+  + \sum_{k,q} \p{h_{kq} v_q - h_k v_k} V_k \ket\psi.
 \end{align}
-Denoting a vector of the coefficients $v_q$ by $\v v$, a matrix of all
-coupling constants $h_{pq}$ by $\m h$, and a vector of all
-$h_q\equiv\sum_k h_{kq}$ by $\v h$, we thus find that the vector
-$\V_1\ket\psi$ is an eigenvector of the interaction Hamiltonian $H_0$
-if the coefficients $\v v$ satisfy
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Two-body perturbation}
+\label{sec:H_V2_psi}
+
+We wish to simplify
 \begin{align}
-  \m h \c \v v - \v h_\d\oc\v v = \Delta\p{\v v} \v v,
-  \label{eq:v_cond_1}
+  H_0 \V_2 \ket\psi
+  = \f12 \sum_{\substack{k,\ell\\p<q}} w_{k\ell} h_{pq}
+  \Pi_{pq} X_k Y_\ell \ket\psi,
+  \label{eq:H_V2_psi_start}
 \end{align}
-where $\m h\c\v v$ is an ordinary matrix-vector product of $\m h$ and
-$\v v$, with components $\sum_q h_{kq} v_q$; $\v h_\d\oc\v v$ is an
-element-wise (Hadamard) product of $\v h_\d$ and $\v v$, with
-components $h_k v_k$; and $\Delta\p{\v v}$ is some constant that may
-depend on $\v v$.  If the interaction Hamiltonian $H_0$ is
-translationally invariant, then all $h_q=h$ and
-$\v h_\d\oc\v v=h\v v$, so the condition in \eqref{eq:v_cond_1}
-simplifies to
+which has terms with $p,q\notin\set{k,\ell}$, one term with
+$\set{p,q}=\set{k,\ell}$, and terms with $p,q\in\set{k,\ell}$.  The
+permutation operator $\Pi_{pq}$ acts trivially on
+$X_k Y_\ell \ket\psi$ when $p,q\notin\set{k,\ell}$ or
+$\set{p,q}=\set{k,\ell}$, leaving a sum over $p,q$ of the form
 \begin{align}
-  \m h \c\v v = \sp{\Delta\p{\v v}+h} \v v.
+  \sum_{\substack{p<q\\p,q\notin\set{k,\ell}}} h_{pq} + h_{k\ell}
+  = \sum_{p<q} h_{pq}
+  - \sum_{\substack{p\in\set{k,\ell}\\q\notin\set{k,\ell}}} h_{pq}
+  = E_0 + 2 h_{k\ell} - h_k - h_\ell.
 \end{align}
-If we construct a perturbation $\V_1\p{\v v}$ with coefficients $\v v$
-that satisfy \eqref{eq:v_cond_1}, then
+The corresponding terms in \eqref{eq:H_V2_psi_start} with
+$p,q\notin\set{k,\ell}$ and $\set{p,q}=\set{k,\ell}$ are then
 \begin{align}
-  H_0 \V_1\p{\v v} \ket\psi
-  = \sp{E_0 + \Delta\p{\v v}} \V_1\p{\v v} \ket\psi,
+  E_0 \V_2 \ket\psi + \f12 \sum_{k,\ell}
+  \p{2 h_{k\ell} - h_k - h_\ell}
+  w_{k\ell} X_k Y_\ell \ket\psi,
 \end{align}
-for any fully symmetric state $\ket\psi\in\M_0$, which for
-$\Delta\p{\v v}>0$ implies that
+while the terms in \eqref{eq:H_V2_psi_start} with only one of
+$p,q\in\set{k,\ell}$ simplify to
 \begin{align}
-  \E \V_1\p{\v v} \P_0
-  = \sum_{k>0} \f1{\Delta_k} \P_k \V_1\p{\v v} \P_0
-  = \f1{\Delta\p{\v v}} \V_1\p{\v v} \P_0,
+  \f12 \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}} w_{k\ell}
+  \p{h_{kq} X_q Y_\ell + h_{\ell q} X_k Y_q} \ket\psi
+  = \sum_{k,\ell,q} w_{k\ell} h_{kq} X_q Y_\ell \ket\psi
+  - \sum_{k,\ell} w_{k\ell} h_{k\ell} Z_k \ket\psi,
 \end{align}
-and in turn
+where
 \begin{align}
-  H_1^{(2)}
-  = -\P_0 \V_1\p{\v v} \E \V_1\p{\v v} \P_0
-  = -\f1{\Delta\p{\v v}} \P_0 \V_1\p{\v v}^2 \P_0.
+  Z_k \equiv \f12 \sp{X_k, Y_k}_+.
 \end{align}
-Computing the second-order effective Hamiltonian $H_1^{(2)}$ thus
-requires us to simplify the product
+In total,
+\begin{align}
+  H_0 \V_2 \ket\psi
+  = E_0 \V_2 \ket\psi
+  + \sum_{k,\ell} \p{\sp{h_{k\ell}-\p{\f{h_k+h_\ell}{2}}} w_{k\ell}
+    + \sum_q h_{kq} w_{q\ell}} X_k Y_\ell \ket\psi
+  - \sum_{k,\ell} h_{k\ell} w_{k\ell} Z_k \ket\psi.
+\end{align}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Simplifying operator products restricted to the
+  fully-symmetric manifold}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Single-body product}
+\label{sec:PVVP}
+
+Computing the second-order effective Hamiltonian $H_1^{(2)}$ induced
+on the fully symmetric manifold $\M_0$ by the perturbation $\V_1$ in
+\eqref{eq:perturbations} requires us to simplify the product
 \begin{align}
@@
   = \sum_p v_p^2 \P_0 V_p^2 \P_0
-  + \sum_{p\ne q} v_p v_q \P_0 V_p V_q \P_0.
+  + \sum_{p\ne q} v_p v_q \P_0 V_p V_q \P_0,
   \label{eq:PVVP_start}
 \end{align}
-We note that vectors $\v v$ satisfying \eqref{eq:v_cond_1} and
-$\Delta\p{\v v}\ne 0$ must be mean-zero, as evident by summing over
-the components of \eqref{eq:v_cond_1}:
-\begin{align}
-  \Delta\p{\v v} \sum_k v_k
-  = \sum_{k,q} h_{kq} v_q - \sum_k h_k v_k
-  = \sum_q h_q v_q - \sum_k h_k v_k
-  = 0.
-  \label{eq:D_v_0}
-\end{align}
-It follows that the first sum on the right of \eqref{eq:PVVP_start} is
+where the coefficients $\v v$ satisfy \eqref{eq:cond_1} with
+eigenvalue $\Delta\ne0$, which implies that $\v v$ is mean-zero.  The
+first sum on the right of \eqref{eq:PVVP_start} is then
 \begin{align}
@@
   = \var\p{\v v} \P_0 W \P_0,
-  &&
-  W \equiv \sum_p V_p^2,
+  \label{eq:PVVP_eq}
 \end{align}
@@
 \begin{align}
-  \var\p{\v x} \equiv \EE_p\sp{\p{x_p-\EE_q\sp{x_q}}^2}
-  = \EE_p\sp{x_p^2} - \EE_q\sp{x_q}^2
+  \var\p{\v v} \equiv \EE_p\sp{\p{v_p-\EE_q\sp{v_q}}^2}
+  = \EE_p\sp{v_p^2} - \EE_q\sp{v_q}^2,
+  &&
+  W \equiv \sum_p V_p^2.
 \end{align}
-is the variance of the elements in the vector $\v x$.  Similarly, the
-second sum on the right of \eqref{eq:PVVP_start} is
+The second sum in \eqref{eq:PVVP_start} is
 \begin{align}
@@
 \end{align}
-where if the coefficients in $\v v$ are mean-zero, then
+where the sum over coefficients in \eqref{eq:PVVP_neq_start} can be
+simplified to
 \begin{align}
   \sum_{p\ne q} v_p v_q
-  = \sum_p v_p \sum_{q\ne p} v_q
-  = \sum_p v_p \p{\sum_q v_q - v_p}
-  = - \sum_p v_p^2
-  = - N \var\p{\v v}.
+  = \sum_{p,q} v_p v_q - \sum_{p=q} v_p v_q
+  = - \sum_p v_p^2 = - N \var\p{\v v}.
 \end{align}
-To evaluate the operator content of \eqref{eq:PVVP_neq_start}, we
-expand
+In order to simplify the product of operators in
+\eqref{eq:PVVP_neq_start}, we expand
 \begin{align}
@@
   \sum_{p\ne q} v_p v_q \P_0 V_p V_q \P_0
-  = -\f{\var\p{\v v}}{N-1}\P_0\p{V^2 - W} \P_0.
+  = - \var\p{\v v} N \P_0 V_0 V_1 \P_0
+  = -\f{\var\p{\v v}}{N-1} \P_0 \p{V^2 - W} \P_0
   \label{eq:PVVP_neq}
@@
 \end{align}
-and so
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Two-body product at first order}
+\label{sec:PXYP}
+
+Computing the first-order effective Hamiltonian $H_2^{(1)}$ induced on
+the fully symmetric manifold $\M_0$ by the perturbation $\V_2$ in
+\eqref{eq:perturbations} requires us to simplify the product
 \begin{align}
-  H_1^{(2)} = \f{\var\p{\v v}}{\p{N-1}\Delta\p{\v v}} \p{V^2 - N W},
+  \P_0 \V_2 \P_0
+  = \sum_{p<q} w_{pq} \P_0 X_p Y_q \P_0
+  = w \times N \P_0 X_0 Y_1 \P_0,
+  \label{eq:PXYP}
 \end{align}
-where we implicitly restrict the effective Hamiltonian $H_1^{(2)}$ to
-the ground-state manifold $\M_0$, and note that the result above holds
-only for perturbations $\V_1\p{\v v}$ constructed from vectors $\v v$
-that satisfy the eigenvalue equation in \eqref{eq:v_cond_1}.
+where
+\begin{align}
+  w \equiv \f1N \sum_{p<q}
+\end{align}
+is the mean value of the pair-wise coefficients in $\V_2$.  To
+simplify the operator content of \eqref{eq:PXYP}, we expand
+\begin{align}
+  \f12 \P_0 \sp{X, Y}_+ \P_0
+  = \P_0 Z \P_0 + N\p{N-1} \P_0 X_0 Y_1 \P_0,
+\end{align}
+where $\sp{A,B}_+\equiv AB+BA$ is an anti-commutator and
+\begin{align}
+  Z \equiv \f12 \sum_p \sp{X_p,Y_p}_+.
+\end{align}
+We thus find that
+\begin{align}
+  \P_0 \V_2 \P_0
+  = \f{w}{N-1} \P_0 \p{\f12 \sp{X,Y}_+ -  Z } \P_0.
+\end{align}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Two-body perturbations}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Two-body product at second order}
+\label{sec:PVVP_2}
 
-The first-order effective Hamiltonian $H_2^{(1)}$ induced on the fully
-symmetric manifold $\M_0$ by the two-body perturbation in $\V_2$ in
-\eqref{eq:perturbations} is
+Computing the second-order effective Hamiltonian $H_2^{(2)}$ induced
+on the fully symmetric manifold $\M_0$ by the perturbation $\V_2$ in
+\eqref{eq:perturbations} requires us to simplify the product
 \begin{align}
-  H_2^{(1)}
-  = \P_0 \V_2 \P_0
-  = \sum_{p<q} v_{pq} \P_0 V_p V_q \P_0
-  \simeq \f{v}{N-1} \p{V^2 - W},
-  &&
-  v \equiv \f1N \sum_{p<q} v_{pq},
+  \P_0 \V_2\p{\m w}^2 \P_0
+  = \f14 \sum_{k,\ell,p,q} w_{k\ell} w_{pq} \P_0 X_k Y_\ell X_p Y_q \P_0.
 \end{align}
-where we made use of the previous result in \eqref{eq:PVVP_neq_ops}.
-In order to compute the second-order effective Hamiltonian
-$H_2^{(2)}$, as before we pick an arbitrary fully symmetric state
-$\ket\psi\in\M_0$ and expand
+The terms with $p,q\notin\set{k,\ell}$ are
 \begin{align}
-  H \V_2 \ket\psi
-  = \sum_{\substack{k<\ell\\p<q}} v_{k\ell} h_{pq}
-  \Pi_{pq} V_k V_\ell \ket\psi
-  = \f12 \sum_{\substack{k,\ell\\p<q}} v_{k\ell} h_{pq}
-  \Pi_{pq} V_k V_\ell \ket\psi.
-  \label{eq:H_V2_psi_start}
+  \f14 \sum_{\substack{k,\ell\\p,q\notin\set{k,\ell}}} w_{k\ell} w_{pq}
+  \P_0 X_0 Y_1 X_2 Y_3 \P_0.
 \end{align}
-The permutation operator $\Pi_{pq}$ acts trivially on
-$V_k V_\ell \ket\psi$ when $p,q\notin\set{k,\ell}$ or
-$\set{p,q}=\set{k,\ell}$, leaving a sum over $p,q$ of the form
-\begin{align}
-  \sum_{\substack{p<q\\p,q\notin\set{k,\ell}}} h_{pq} + h_{k\ell}
-  = \sum_{p<q} h_{pq}
-  - \sum_{\substack{p<q\\p\in\set{k,\ell}\\q\notin\set{k,\ell}}} h_{pq}
-  - \sum_{\substack{p<q\\p\notin\set{k,\ell}\\q\in\set{k,\ell}}} h_{pq}
-  = E_0 - \sum_{\substack{p\in\set{k,\ell}\\q\notin\set{k,\ell}}} h_{pq}
-  = E_0 + 2 h_{k\ell} - h_k - h_\ell.
-\end{align}
-The corresponding terms in \eqref{eq:H_V2_psi_start} with
-$p,q\notin\set{k,\ell}$ and $\set{p,q}=\set{k,\ell}$ are then
-\begin{align}
-  E_0 \V_2 \ket\psi + \f12 \sum_{k,\ell}
-  \p{2 h_{k\ell} - h_k - h_\ell} v_{k\ell} V_k V_\ell \ket\psi,
-\end{align}
-while the terms in \eqref{eq:H_V2_psi_start} with only one of
-$p,q\in\set{k,\ell}$ simplify to
-\begin{align}
-  \f12 \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}}
-  v_{k\ell} \p{h_{kq} V_q V_\ell + h_{\ell q} V_k V_q} \ket\psi
-  = \sum_{k,\ell,q} v_{k\ell} h_{kq} V_q V_\ell \ket\psi
-  - \sum_{k,\ell} v_{k\ell} h_{k\ell} V_k^2 \ket\psi.
-\end{align}
-In total,
-\begin{align}
-  H_0 \V_2 \ket\psi
-  = E_0 \V_2 \ket\psi
-  + \sum_{k,\ell} \p{\sp{h_{k\ell}-\p{\f{h_k+h_\ell}{2}}} v_{k\ell}
-    + \sum_q h_{kq} v_{q\ell}} V_k V_\ell \ket\psi
-  - \sum_{k,\ell} h_{k\ell} v_{k\ell} V_k^2 \ket\psi.
-  \label{eq:H_V2_psi_mid}
-\end{align}
-We thus find that the vector $\V_2\ket\psi$ is an eigenvector of the
-interaction Hamiltonian $H_0$ if the matrix $\m v$ of all $v_{k\ell}$
-satisfies, for some constant $\e\p{\m v}$ that may depend on $\m v$,
-\begin{align}
-  \bar{\m h}\oc\m v + \m h\c\m v = \e\p{\m v} \m v,
-  &&
-  \bar{\m h}
-  \equiv \m h - \f12\p{\v h_\d\otimes\v\1 + \v\1\otimes\v h_\d},
-  \label{eq:v_cond_2}
-\end{align}
-where $\bar{\m h}\oc\m v$ denotes an element-wise (Hadamard) product
-of $\bar{\m h}$ and $\m v$, with components
-$\tilde h_{k\ell} v_{k\ell}$; $\m h\c\m v$ denotes an ordinary matrix
-product of $\m h$ and $\m v$, with components
-$\sum_q h_{kq} v_{q\ell}$; $\v\1$ is a constant vector of ones; and
-$\v h_\d\otimes\v\1$ is a tensor product of $\v h_\d$ and $\v\1$.
-Note that the diagonal components of the condition in
-\eqref{eq:v_cond_2} automatically enforce that the last sum in
-\eqref{eq:H_V2_psi_mid} vanishes.  If the interaction Hamiltonian
-$H_0$ is translationally invariant, then all $h_q=h$ and
-$\v h_\d\otimes\v\1\oc\m v=\v\1\otimes\v h_\d\oc\m v=h\m v$, so the
-condition in \eqref{eq:v_cond_2} becomes
-\begin{align}
-  \m h\oc\m v + \m h\c\m v = \sp{\e\p{\m v}+h} \m v,
-\end{align}
-If we construct a perturbation $\V_2\p{\m v}$ with coefficients $\m v$
-that satisfy \eqref{eq:v_cond_2}, then
-\begin{align}
-  H_0 \V_2\p{\m v} \ket\psi
-  = \sp{E_0 + 2\e\p{\m v}} \V_2\p{\m v} \ket\psi.
-\end{align}
-As a curiosity, we note that if $\V_1\p{\v v}\ket\psi$ is an
-eigenvector of $H_0$, then the coefficient vector $\v v$ must satisfy
-\eqref{eq:D_v_0}, which we used to determine that $\v v$ was mean-zero
-if $\Delta\p{\v v}\ne0$; the analogous condition on matrices $\m v$ in
-the two-body case is
-\begin{align}
-  N \e\p{\m v} v = \f12 \sum_{k,\ell} h_{k\ell} v_{k\ell},
-\end{align}
-where $v\equiv\sum_{p<q}v_{pq}/N$ is the mean value of the
-coefficients in $\V_2\p{\m v}$.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Constructing eigenstates of SU($n$)-symmetric interactions}
-
-Here we discuss the construction of eigenstates of the interaction
-Hamiltonian $H_0$.  If a vector $\v v$ and matrix $\m v$ satisfy the
-conditions in \eqref{eq:v_cond_1} and \eqref{eq:v_cond_2}, then the
-single- and two-body operators $\V_1\p{\v v}$ and $\V_2\p{\m v}$
-generate eigenvectors of the interaction Hamiltonian $H_0$ when they
-act on any fully symmetric state $\ket\psi\in\M_0$.  We can use this
-fact to construct eigenstates of $H_0$ by finding solutions to
-\eqref{eq:v_cond_1} and \eqref{eq:v_cond_2}.
-
-The single-body case is straightforward: finding vectors $\v v$ that
-satisfy \eqref{eq:v_cond_1} is equivalent to finding eigenvectors of
-the matrix $\m h-\m h_\d$, where $\m h_\d\equiv\diag\v h_\d$ is a
-diagonal matrix.  The constant $\Delta\p{\v v}$ is then the eigenvalue
-of $\m h-\m h_\d$ that corresponds to the eigenvector $\v v$.  The
-task of constructing eigenvectors $\V_1\p{\v v}\ket\psi$ of $H_0$ and
-determining their energy thus reduces to finding the eigenvectors and
-eigenvalues of the $N\times N$ matrix $\m h-\m h_\d$.
-
-The two-body case requires a bit of work to re-cast as an
-eigenvector/eigenvalue problem.  We first write the matrix $\m v$ as a
-vector,
-\begin{align}
-  \m v = \sum_{k,\ell} v_{k\ell} \op{k}{\ell}
-  \to \v{\m v} \equiv \sum_{k,\ell} v_{k\ell} \ket{k\ell},
-\end{align}
-and define the matrix
-\begin{align}
-  \check{\m h}
-  \equiv \sum_{k,\ell} \bar h_{k\ell} \op{k\ell}
-  + \sum_{k,q,\ell} h_{kq} \op{k\ell}{q\ell},
-  \label{eq:h_super_mat}
-\end{align}
-in terms of which the condition in \eqref{eq:v_cond_2} becomes
-\begin{align}
-  \check{\m h} \c \v{\m v} = \e\p{\m v} \v{\m v}.
-\end{align}
-Note that the matrix $\check{\m h}$ in \eqref{eq:h_super_mat} can be
-written in the block-diagonal form
-\begin{align}
-  \check{\m h} = \sum_\ell \m h_\ell \otimes \op{\ell},
-  &&
-  \m h_\ell
-  \equiv \sum_{k,q}\p{\delta_{kq} \bar h_{k\ell} + h_{kq}} \op{k}{q},
-\end{align}
-where each block $\m h_\ell$ is an $N\times N$ matrix.  We can
-diagonalize these blocks independently, and classify their
-eigenvectors by the corresponding eigenvalues.  ``Total'' eigenvectors
-$\v{\m v}$ of $\check{\m h}$ are then essentially built from
-eigenvectors within each block $\m h_\ell$ that share the same
-eigenvalue.  The task of constructing eigenvectors
-$\V_2\p{\m v}\ket\psi$ of $H_0$ and determining their energy thus
-reduces to finding the eigenvectors and eigenvalues of $N$ matrices
-$\m h_\ell$ of size $N\times N$.
-
-If the interaction Hamiltonian $H_0$ is translationally invariant,
-then $\bar h_{k\ell}=h_{k\ell}-h$, and $h_{k\ell}$ depends only on the
-displacement between the spins indexed by $k$ and $\ell$.  The blocks
-$\m h_\ell$ then take the form
-\begin{align}
-  \m h_{\ell} = \m\omega_\ell \c \m h_0 \c \m\omega_\ell^\dag,
-  &&
-  \m\omega_\ell \equiv \sum_k \op{k+\ell}{k},
-  \label{eq:block_shift}
-\end{align}
-which implies that we need only find the eigenvectors and eigenvalues
-of a single block $\m h_0$, after which we can determine the
-eigenvectors of any block simply by appropriate applications of the
-shift matrix $\m\omega_\ell$.  Note that in dimensions $D>1$, the
-relation in \eqref{eq:block_shift} holds for vectors $k,\ell$ that
-index lattice sites, i.e.~$k=\p{k_1,k_2,\cdots,k_D}$ and likewise with
-$\ell$.
+\red{[to do: finish calculation]}
 

### 2019-11-01 collective_sun/sun_perturbation.tex
major improvements to notation

@@
   \Pi_{pq} \equiv \sum_{\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)},
+  \label{eq:ints}
 \end{align}
@@
 $p$ to $\ket\mu$ from $\ket\nu$; the operator $\Pi_{pq}$ permutes
-spins $p$ and $q$; and the coefficients $h_{pq}$ satisfy
-\begin{align}
-  h_{pp} = 0
-  &&
-  h_{pq} \le 0,
-  &&
-  h_{pq} = h_{qp}.
-\end{align}
+spins $p$ and $q$; and the $h_{pq}$ in the Hamiltonian are all
+negative: $h_{pq}<0$.
+
 The ground-state manifold $\M_0$ of the interaction Hamiltonian $H_0$
@@
 throughout these notes that the interaction Hamiltonian $H_0$ is
-gapped, such that there is a finite energy difference between $\M_0$
+gapped, i.e.~that there is a finite energy difference between $\M_0$
 and its orthogonal complement.  The energy of fully symmetric states
@@
 \begin{align}
-  E_0 \equiv \sum_{p<q} h_{pq} = \f12 \sum_{p,q} h_{pq}
-  = \f12 \sum_p h_p = \f12 N h,
+  E_0 \equiv \sum_{p<q} h_{pq}
+  = \f12 \sum_{p,q} h_{pq}
+  = \f12 \sum_p h_p
+  = \f12 N h,
 \end{align}
-where we define
+where for convenience define coefficients $h_{pq}$ for all $p,q$ (as
+opposed to only $p<q$) by enforcing $h_{pq}=h_{qp}$ and $h_{pp}=0$;
+and
 \begin{align}
@@
   &&
-  h \equiv \EE_p\sp{h_p} = \f1N \sum_{p,q} h_{pq},
-  &&
-  \EE_p\sp{X_p} \equiv \f1N \sum_p X_p.
+  h \equiv \EE_p\sp{h_p} = \f1N \sum_p h_p.
 \end{align}
@@
 \begin{align}
-  \V_1 \equiv \sum_p v_p V_p,
+  \V_1 \equiv \sum_p v_p V_p = \v v\c\v V,
   &&
   \V_2 \equiv \sum_{p<q} w_{pq} X_p Y_q
-  = \f12 \sum_{p,q} w_{pq} X_p Y_q,
+  = \f12 \sum_{p,q} w_{pq} X_p Y_q
+  = \f12 \v X \c\m w \c\v Y,
   \label{eq:perturbations}
 \end{align}
-where $V_p,X_p,Y_p$ are trace-zero operators on spin $p$, and we
-enforce $w_{pq}=w_{qp}$ with $w_{pp}=0$.  The effective Hamiltonian
-$H_M$ induced on the ground-state manifold $\M_0$ by an $M$-body
-perturbation $\V_M$ through second order in perturbation theory is
-given by\cite{bravyi2011schrieffer, perlin2019effective}
+where $V_p,X_p,Y_p$ are trace-zero operators on spin $p$;
+$\v V,\v X,\v Y$ are respectively vectors of all $V_p,X_p,Y_p$; $\v v$
+and $\m w$ are respectively a vector and matrix of all $v_p,w_{pq}$;
+we enforce $w_{pq}=w_{qp}$ and $w_{pp}=0$; and we use a dot ($\c$) to
+denote an inner product between matrices or vectors as appropriate.
+
+The effective Hamiltonian $H_M$ induced on the ground-state manifold
+$\M_0$ by an $M$-body perturbation $\V_M$ through second order in
+perturbation theory is given by\cite{bravyi2011schrieffer,
+  perlin2019effective}
 \begin{align}
@@
 \begin{align}
-  H_1^{(1)} = \sum_p v_p \P_0 V_p \P_0,
+  H_1^{(1)} = \v v \c \P_0 \v V \P_0,
 \end{align}
@@
   &&
-  V = \sum_p V_p,
+  V \equiv \sum_p V_p,
   \label{eq:H_1_1}
@@
   = E_0 \V_1 \ket\psi
-  + \sum_{k,q} \p{h_{kq} v_q - h_k v_k} V_k \ket\psi.
-  \label{eq:H_V1_psi}
+  + \p{\v v\c\m h - \v v\oc\v h}\c \v V \ket\psi,
+  \label{eq:H_V1_psi_start}
 \end{align}
-Denoting a vector of the coefficients $v_q$ by $\v v$, a matrix of all
-coupling constants $h_{pq}$ by $\m h$, and a vector of all
-$h_q\equiv\sum_k h_{kq}$ by $\v h$, we thus find that the vector
+where $\m h$ is a matrix of all coefficients $h_{pq}$ in the
+interaction Hamiltonian $H_0$ in \eqref{eq:ints}, and $\v v\oc\v h$ is
+the element-wise (Hadamard) product of $\v v$ and $\v h$, i.e.~a
+vector with components $v_kh_k$.  We thus find that the vector
 $\V_1\ket\psi$ is an eigenvector of the interaction Hamiltonian $H_0$
@@
   \m h \c \v v - \v h\oc\v v
-  = \p{\m h-\diag\v h}\c\v v
+  = \p{\m h - \diag\v h}\c\v v
   = \Delta\p{\v v} \v v,
@@
 \end{align}
-where $\m h\c\v v$ is an ordinary matrix-vector product of $\m h$ and
-$\v v$, with components $\sum_q h_{kq} v_q$; $\v h\oc\v v$ is an
-element-wise (Hadamard) product of $\v h$ and $\v v$, with components
-$h_k v_k$; $\diag\v h$ is a matrix with $\v h$ on the diagonal and
-zeroes everywhere else; and $\Delta\p{\v v}$ is some constant that may
-depend on $\v v$.  If the interaction Hamiltonian $H_0$ is
-translationally invariant, then all $h_q=h$ and $\v h\oc\v v=h\v v$,
-so the condition in \eqref{eq:cond_1} simplifies to
+where $\diag\v h$ is a matrix with $\v h$ on the diagonal and zeroes
+everywhere else; and $\Delta\p{\v v}$ is some constant that may depend
+on $\v v$.  If the interaction Hamiltonian $H_0$ is translationally
+invariant, then all $h_q=h$ and $\v h\oc\v v=h\v v$, so the condition
+in \eqref{eq:cond_1} simplifies to
 \begin{align}
@@
   = \sp{E_0 + \Delta} \V_1\p{\v v_\Delta} \ket\psi.
+  \label{eq:H_V1_psi}
 \end{align}
+Interestingly, the energy of the (unnormalized) state
+$\V_1\p{\v v}\ket\psi$ depends only on the coefficients $\v v$, and is
+entirely independent of the state $\ket\psi$ or choice of single-spin
+trace-zero operator used to build $\V_1$ in \eqref{eq:perturbations}.
 Finding operators $\V_1\p{\v v}$ that generate eigenvectors of the
 interaction Hamiltonian $H_0$ when they act on fully symmetric states
-$\ket\psi$ thus reduces to finding eigenvectors $\v v$ of
-$\m h-\diag\v h$.  Interestingly, the energy of the (unnormalized)
-state $\V_1\p{\v v}\ket\psi$ depends only on the coefficients $\v v$,
-and is entirely independent of the state $\ket\psi$ or choice of
-single-spin trace-zero operator used to build $\V_1$ in
-\eqref{eq:perturbations}.
+$\ket\psi$ thus reduces to finding eigenvectors $\v v$ of the
+interaction matrix $\m h-\diag\v h$.
 
 If the interaction Hamiltonian $H_0$ is gapped, then a vector
-$\V_1\p{\v v_\Delta}\ket\psi$ with $\Delta=0$ must lie with in the
+$\V_1\p{\v v_\Delta}\ket\psi$ with $\Delta=0$ must lie within the
 fully symmetric manifold $\M_0$, which implies that the operator
-$\V_1\p{\v v_\Delta}$ preserves the permutational symmetry of
-$\ket\psi$.  The only way for $\V_1\p{\v v_\Delta}$ to obey
-permutational symmetry is for $\v v_\Delta$ to be a constant vector,
+$\V_1\p{\v v_\Delta}$ preserves the permutational symmetry of $\M_0$.
+The only way for $\V_1\p{\v v_\Delta}$ to obey permutational symmetry
+is for $\v v_\Delta$ to be a constant vector,
 i.e.~$\v v_\Delta=v_\Delta\v\1$ with $\1$ a constant vector of ones.
@@
   H_1^{(2)} \simeq \f1{N-1} \sum_{\Delta>0}
-  \f{\var\p{\v v_\Delta}}{\Delta} \p{V^2 - N W},
+  \f{\var\p{\v v_\Delta}}{\Delta} \p{V^2 - N W}.
   \label{eq:H_1_2}
 \end{align}
-where $\simeq$ denotes equality up to a restriction to the fully
+Here $\simeq$ denotes equality up to a restriction to the fully
 symmetric manifold $\M_0$, and
 \begin{align}
-  \var\p{\v x} \equiv \EE_p\sp{\p{x_p-\EE_q\sp{x_q}}^2},
+  \var\p{\v x} \equiv \EE_p\sp{\p{x_p-\EE_q\sp{x_q}}^2}
+  = \EE_p\sp{x_p^2} - \EE_q\sp{x_q}^2,
   &&
-  W \equiv \sum_p V_p^2.
+  W \equiv \sum_p V_p^2,
 \end{align}
+with $\EE_p\sp{x_p}\equiv\sum_px_p/N$ the mean of $\v x$.
 
@@
   H_2^{(1)}
-  = \P_0 \V_2 \P_0
-  = \sum_{p<q} w_{pq} \P_0 X_p Y_q \P_0
-  \simeq \f{w}{N-1} \p{\f12 \sp{X,Y}_+ - Z},
+  = \f12 \P_0 \v X\c\m w\c\v Y\P_0
+  \simeq \f12 w \p{\tilde Z - Z},
 \end{align}
-where
+where in terms of the anti-commutator $\sp{A,B}_+\equiv AB+BA$ and
+collective operators $X\equiv\sum_p X_p$ and $Y\equiv\sum_p Y_p$, we
+define
 \begin{align}
-  w \equiv \f1N \sum_{p<q} w_{pq},
+  w \equiv {N\choose 2}^{-1} \sum_{p<q} w_{pq},
   &&
-  X \equiv \sum_p X_p,
+  \tilde Z \equiv \f12 \sp{X,Y}_+
   &&
-  Y \equiv \sum_p Y_p,
-  &&
-  Z \equiv \f12 \sum_p \sp{X_p, Y_p}_+,
+  Z \equiv \sum_p \f12 \sp{X_p, Y_p}_+.
 \end{align}
-and $\sp{X,Y}_+\equiv XY+YX$ is an anti-commutator.  In order to
-compute the second-order effective Hamiltonian $H_2^{(2)}$, as before
-we pick an arbitrary fully symmetric state $\ket\psi\in\M_0$ and
-expand (see Appendix \ref{sec:H_V2_psi})
+In order to compute the second-order effective Hamiltonian
+$H_2^{(2)}$, as before we pick an arbitrary fully symmetric state
+$\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_V2_psi})
 \begin{align}
@@
   = E_0 \V_2 \ket\psi
-  + \sum_{k,\ell} \p{\sp{h_{k\ell}-\p{\f{h_k+h_\ell}{2}}} w_{k\ell}
-    + \sum_q h_{kq} w_{q\ell}} X_k Y_\ell \ket\psi
-  - \sum_{k,\ell} h_{k\ell} w_{k\ell} Z_k \ket\psi,
+  + \v X \c \p{\bar{\m h}\oc\m w + \m h\c\m w} \c \v Y \ket\psi
+  - \v\1 \c \p{\m h\oc\m w} \c \v Z \ket\psi,
   \label{eq:H_V2_psi}
 \end{align}
-where $Z_k\equiv\p{1/2}\sp{X_k,Y_k}_+$.  We thus find that the vector
-$\V_2\ket\psi$ is an eigenvector of the interaction Hamiltonian $H_0$
-if the matrix $\m w$ of all $w_{k\ell}$ satisfies
+where $\v\1$ is an $N$-component vector of ones,
 \begin{align}
-  \bar{\m h}\oc\m w + \m h\c\m w = \e\p{\m w} \m w,
-  &&
   \bar{\m h}
   \equiv \m h - \f12\p{\v h\otimes\v\1 + \v\1\otimes\v h},
+\end{align}
+is a modified interaction matrix with entries $h_{pq}-\p{h_p+h_q}/2$;
+and $\m x\oc\m y$ is an element-wise (Hadamard) product of $\m x$ and
+$\m y$, with entries $x_{pq}y_{pq}$.  We thus find that the vector
+$\V_2\ket\psi$ is an eigenvector of the interaction Hamiltonian $H_0$
+if the matrix $\m w$ of all $w_{pq}$ satisfies
+\begin{align}
+  \bar{\m h}\oc\m w + \m h\c\m w = \f12 \Delta\p{\m w} \m w,
   \label{eq:cond_2}
 \end{align}
-where $\bar{\m h}\oc\m w$ denotes an element-wise (Hadamard) product
-of $\bar{\m h}$ and $\m w$, with components
-$\bar h_{k\ell} w_{k\ell}$; $\m h\c\m w$ denotes an ordinary matrix
-product of $\m h$ and $\m w$, with components
-$\sum_q h_{kq} w_{q\ell}$; $\e\p{\m w}$ is some constant that may
-depend on $\m w$; $\v\1$ is a constant vector of ones; and
-$\v h\otimes\v\1$ is a tensor product of $\v h$ and $\v\1$.  Note that
+for some constant $\Delta\p{\m w}$ that may depend on $\m w$; $\v\1$
+is a constant vector of ones.  Remembering that both $\m h$ and $\m w$
+are both symmetric (under transposition) with zeros on the diagonal,
 the diagonal components of the condition in \eqref{eq:cond_2}
-automatically enforce that the last sum in \eqref{eq:H_V2_psi}
+automatically enforce that the last term in \eqref{eq:H_V2_psi}
 vanishes.
@@
 \begin{align}
-  \m w = \sum_{k,\ell} w_{k\ell} \op{k}{\ell}
-  \to \v{\m w} \equiv \sum_{k,\ell} w_{k\ell} \ket{k\ell},
+  \m w = \sum_{p,q} w_{pq} \op{p}{q}
+  \to \v{\m w} \equiv \sum_{p,q} w_{pq} \ket{pq},
 \end{align}
@@
   \check{\m h}
-  \equiv \sum_{k,\ell} \bar h_{k\ell} \op{k\ell}
-  + \sum_{k,q,\ell} h_{kq} \op{k\ell}{q\ell},
+  \equiv \sum_{p,q} \bar h_{pq} \op{pq}
+  + \sum_{p,q,\ell} h_{pq} \op{p\ell}{q\ell},
   \label{eq:h_super_mat}
@@
 \begin{align}
-  \check{\m h} \c \v{\m w} = \e\p{\m w} \v{\m w}.
+  \check{\m h} \c \v{\m w} = \f12 \Delta\p{\m w} \v{\m w}.
   \label{eq:cond_2_eig}
@@
   \m h_\ell
-  \equiv \sum_{k,q}\p{\delta_{kq} \bar h_{k\ell} + h_{kq}} \op{k}{q},
+  \equiv \sum_{p,q}\p{\delta_{pq} \bar h_{p\ell} + h_{pq}} \op{p}{q},
 \end{align}
@@
 \begin{align}
-  \m h_\ell \c \v{\m w}_\ell = \e\p{\m w} \v{\m w}_\ell,
+  \m h_\ell \c \v{\m w}_\ell = \f12 \Delta\p{\m w} \v w_\ell,
   &&
-  \v{\m w}_\ell \equiv \sum_k w_{k\ell} \ket{k}.
+  \v w_\ell \equiv \sum_p w_{p\ell} \ket{p}.
   \label{eq:cond_2_eig_block}
@@
 We can thus (numerically) diagonalize the blocks $\m h_\ell$
-independently, and classify their eigenvectors $\v{\m w}_\ell$ by the
-corresponding eigenvalues $\e\p{\m w}$.  ``Total'' eigenvectors
+independently, and classify their eigenvectors $\v w_\ell$ by the
+corresponding eigenvalues $\Delta\p{\m w}/2$.  ``Total'' eigenvectors
 $\v{\m w}$ of $\check{\m h}$ are then built from eigenvectors
 $\v{\m w}_\ell$ within each block $\m h_\ell$ that share the same
-eigenvalue $\e\p{\m w}$.  The task of constructing eigenvectors
+eigenvalue $\Delta\p{\m w}/2$.  The task of constructing eigenvectors
 $\V_2\p{\m w}\ket\psi$ of $H_0$ and determining their energy thus
@@
   &&
-  \m\Sigma_\ell \equiv \sum_k \op{k+\ell}{k},
+  \m\Sigma_\ell \equiv \sum_p \op{p+\ell}{p},
   \label{eq:block_shift}
@@
 shift matrix $\m\Sigma_\ell$.  Note that in dimensions $D>1$, the
-relation in \eqref{eq:block_shift} holds for vectors $k,\ell$ that
-index lattice sites, i.e.~$k=\p{k_1,k_2,\cdots,k_D}$ and likewise with
+relation in \eqref{eq:block_shift} holds for vectors $p,\ell$ that
+index lattice sites, i.e.~$p=\p{p_1,p_2,\cdots,p_D}$ and likewise with
 $\ell$.
 
+If we {\it construct} a perturbation $\V_2\p{\m w_\Delta}$ with
+coefficients $\m w_\Delta$ that satisfy \eqref{eq:cond_1} with
+eigenvalue $\Delta/2$, then
+\begin{align}
+  H_0 \V_2\p{\m w_\Delta} \ket\psi
+  = \sp{E_0 + \Delta} \V_2\p{\m w_\Delta} \ket\psi.
+  \label{eq:H_V1_psi}
+\end{align}
 As with the case of single-body perturbations, we can decompose any
-matrix $\m w$ into its projections $\m w_\e$ onto the eigenspace of
-matrices satisfying \eqref{eq:cond_2} with eigenvalue $\e$,
-i.e.~$\m w=\sum_\e\m w_\e$, which allows us to expand
+matrix $\m w$ into its projections $\m w_\Delta$ onto the eigenspace
+of matrices satisfying \eqref{eq:cond_2} with eigenvalue $\Delta/2$,
+i.e.~$\m w=\sum_\Delta\m w_\Delta$, which allows us to expand
 \begin{align}
-  \V_2\p{\m w} = \sum_\e \V_2\p{\m w_\e}.
+  \V_2\p{\m w} = \sum_\Delta \V_2\p{\m w_\Delta}.
 \end{align}
@@
   H_2^{(2)} = - \P_0 \V_2\p{\m w} \E \V_2\p{\m w} \P_0
-  = -\sum_{\e>0} \f1\e \P_0 \V_2\p{\m w_\e} \P_\e \V_2\p{\m w_\e} \P_0
-  = -\sum_{\e>0} \f1\e \P_0 \V_2\p{\m w_\e}^2 \P_0,
+  = -\sum_{\Delta>0} \f1\Delta \P_0 \V_2\p{\m w_\Delta}^2 \P_0,
 \end{align}
-where the product $\P_0 \V_2\p{\m w_\e}^2 \P_0$ is worked out in
-Appendix \ref{sec:PVVP_2}, giving us \red{[to do: finish calculation]}
-\begin{align}
-  H_2^{(2)} =
-\end{align}
-
-\vspace{3cm}
+where the product $\P_0 \V_2\p{\m w_\Delta}^2 \P_0$ can be written in
+terms collective operators using techniques similar to those in
+Appendix \ref{sec:sym_prod}.
 
@@
 Here we simplify vectors of the form $H_0\V_M\ket\psi$ to arrive at
-the expansions in \eqref{eq:H_V1_psi} and \eqref{eq:H_V2_psi}.
+the expansions in \eqref{eq:H_V1_psi_start} and \eqref{eq:H_V2_psi}.
 
@@
   - \f12 \sum_{\substack{p,q\\k\in\set{p,q}}} h_{pq}
-  = \f12 N h - \sum_q h_{kq}
   = E_0 - h_k,
 \end{align}
-where we have used the facts that $h_{pp}=0$ and $h_{pq}=h_{qp}$.  The
+where we have used the facts that $h_{pq}=h_{qp}$ and $h_{pp}=0$.  The
 terms with $k\notin\set{p,q}$ in \eqref{eq:H_V1_psi_start} are then
@@
   = \sum_k \p{E_0 - h_k} v_k V_k \ket\psi
-  = E_0 \V_1 \ket\psi - \sum_k h_k v_k V_k \ket\psi.
+  = E_0 \V_1 \ket\psi - \v h\oc\v v \c\v V \ket\psi,
 \end{align}
-In the case of $k\in\set{p,q}$, we have
+where $\v h\oc\v v$ is an element-wise (Hadamard) product of $\v h$
+and $\v v$.  In the case of $k\in\set{p,q}$, we have
 \begin{align}
@@
   + \sum_{p<k} h_{pk} v_k V_p \ket\psi
-  = \sum_{k,q} h_{kq} v_k V_q \ket\psi,
+  = \sum_{k,q} h_{kq} v_k V_q \ket\psi
+  = \v v\c\m h\c\v V \ket\psi,
 \end{align}
-which with some re-indexing gives us
+which implies
 \begin{align}
@@
   = E_0 \V_1 \ket\psi
-  + \sum_{k,q} \p{h_{kq} v_q - h_k v_k} V_k \ket\psi.
+  + \p{\v v\c\m h - \v v\oc\v h}\c \v V \ket\psi.
 \end{align}
@@
 \begin{align}
-  E_0 \V_2 \ket\psi + \f12 \sum_{k,\ell}
-  \p{2 h_{k\ell} - h_k - h_\ell}
-  w_{k\ell} X_k Y_\ell \ket\psi,
+  E_0 \V_2 \ket\psi + \sum_{k,\ell}
+  \sp{h_{k\ell} - \f12\p{h_k+h_\ell}} w_{k\ell} X_k Y_\ell \ket\psi
+  = E_0 \V_2 \ket\psi + \v X \c \p{\bar{\m h}\oc\m w} \c \v Y \ket\psi,
 \end{align}
-while the terms in \eqref{eq:H_V2_psi_start} with only one of
-$p,q\in\set{k,\ell}$ simplify to
+where $\bar{\m h}\oc\m w$ is an element-wise (Hadamard) product
+between $\bar{\m h}$ and $\m w$,
+\begin{align}
+  \bar{\m h} \equiv \m h - \f12\p{\v h\otimes\v\1 + \v\1\otimes\v h},
+\end{align}
+and $\v\1$ is a vector of ones.  The terms in
+\eqref{eq:H_V2_psi_start} with only one of $p,q\in\set{k,\ell}$
+simplify to
 \begin{align}
@@
   \p{h_{kq} X_q Y_\ell + h_{\ell q} X_k Y_q} \ket\psi
-  = \sum_{k,\ell,q} w_{k\ell} h_{kq} X_q Y_\ell \ket\psi
-  - \sum_{k,\ell} w_{k\ell} h_{k\ell} Z_k \ket\psi,
+  &= \sum_{k,\ell,q} w_{k\ell} h_{kq} X_q Y_\ell \ket\psi
+  - \sum_{k,\ell} w_{k\ell} h_{k\ell} Z_k \ket\psi \\
+  &= \v X \c\m h \c\m w\c\v Y \ket\psi
+  - \v\1 \c \p{\m h\oc\m w}\c \v Z \ket\psi,
 \end{align}
@@
 \begin{align}
-  Z_k \equiv \f12 \sp{X_k, Y_k}_+.
+  Z_k \equiv \f12 \sp{X_k, Y_k}_+,
 \end{align}
-In total,
+and $\v Z$ is a vector of all $Z_p$.  In total,
 \begin{align}
@@
   = E_0 \V_2 \ket\psi
-  + \sum_{k,\ell} \p{\sp{h_{k\ell}-\p{\f{h_k+h_\ell}{2}}} w_{k\ell}
-    + \sum_q h_{kq} w_{q\ell}} X_k Y_\ell \ket\psi
-  - \sum_{k,\ell} h_{k\ell} w_{k\ell} Z_k \ket\psi.
+  + \v X \c \p{\bar{\m h}\oc\m w + \m h\c\m w} \c \v Y \ket\psi
+  - \v\1 \c \p{\m h\oc\m w}\c \v Z \ket\psi.
 \end{align}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Simplifying operator products restricted to the
-  fully-symmetric manifold}
+\section{Simplifying operator products restricted to the fully
+  symmetric manifold}
+\label{sec:sym_prod}
 
@@
 
-Computing the second-order effective Hamiltonian $H_1^{(2)}$ induced
+Calculating the second-order effective Hamiltonian $H_1^{(2)}$ induced
 on the fully symmetric manifold $\M_0$ by the perturbation $\V_1$ in
@@
 
-Computing the first-order effective Hamiltonian $H_2^{(1)}$ induced on
-the fully symmetric manifold $\M_0$ by the perturbation $\V_2$ in
+Calculating the first-order effective Hamiltonian $H_2^{(1)}$ induced
+on the fully symmetric manifold $\M_0$ by the perturbation $\V_2$ in
 \eqref{eq:perturbations} requires us to simplify the product
@@
   \P_0 \V_2 \P_0
-  = \sum_{p<q} w_{pq} \P_0 X_p Y_q \P_0
-  = w \times N \P_0 X_0 Y_1 \P_0,
-  \label{eq:PXYP}
+  = \f12 \P_0 \v X\c\m w\c\v Y\P_0
+  = \f12 \sum_{p\ne q} w_{pq} \P_0 X_p Y_q \P_0
+  = {N\choose 2} w \P_0 X_0 Y_1 \P_0,
+  \label{eq:PXYP_start}
 \end{align}
-where
+where ${N\choose2}=N\p{N-1}/2$ and
 \begin{align}
-  w \equiv \f1N \sum_{p<q}
+  w \equiv {N\choose 2}^{-1} \sum_{p<q} w_{pq}
 \end{align}
 is the mean value of the pair-wise coefficients in $\V_2$.  To
-simplify the operator content of \eqref{eq:PXYP}, we expand
+simplify the operator content of \eqref{eq:PXYP_start}, we define
 \begin{align}
-  \f12 \P_0 \sp{X, Y}_+ \P_0
-  = \P_0 Z \P_0 + N\p{N-1} \P_0 X_0 Y_1 \P_0,
+  Z_p \equiv \f12\sp{X_p, Y_p}_+,
+  &&
+  Z \equiv \sum_p Z_p,
+  &&
+  \tilde Z \equiv \f12 \sp{X, Y}_+,
 \end{align}
-where $\sp{A,B}_+\equiv AB+BA$ is an anti-commutator and
+and expand
 \begin{align}
-  Z \equiv \f12 \sum_p \sp{X_p,Y_p}_+.
+  \P_0 \tilde Z \P_0 = \P_0 Z \P_0 + N\p{N-1} \P_0 X_0 Y_1 \P_0.
 \end{align}
-We thus find that
+Substituting the product $\P_0 X_0 Y_1 \P_0$ back into
+\eqref{eq:PXYP_start}, we thus find that
 \begin{align}
   \P_0 \V_2 \P_0
-  = \f{w}{N-1} \P_0 \p{\f12 \sp{X,Y}_+ -  Z } \P_0.
+  = \f12 w \P_0 \p{\tilde Z -  Z } \P_0.
 \end{align}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Two-body product at second order}
-\label{sec:PVVP_2}
-
-Computing the second-order effective Hamiltonian $H_2^{(2)}$ induced
-on the fully symmetric manifold $\M_0$ by the perturbation $\V_2$ in
-\eqref{eq:perturbations} requires us to simplify the product
-\begin{align}
-  \P_0 \V_2\p{\m w}^2 \P_0
-  = \f14 \sum_{k,\ell,p,q} w_{k\ell} w_{pq} \P_0 X_k Y_\ell X_p Y_q \P_0.
-\end{align}
-The terms with $p,q\notin\set{k,\ell}$ are
-\begin{align}
-  \f14 \sum_{\substack{k,\ell\\p,q\notin\set{k,\ell}}} w_{k\ell} w_{pq}
-  \P_0 X_0 Y_1 X_2 Y_3 \P_0.
-\end{align}
-
-\red{[to do: finish calculation]}
 

### 2019-11-01 collective_sun/sun_perturbation.tex
some rearranging text

@@
 \begin{align}
-  H_0 = \sum_{p<q} h_{pq} \Pi_{pq},
+  H_0 = \sum_{p<q} h_{pq} \Pi_{pq}
+  = \f12 \sum_{p,q} \Pi_{pq},
   &&
@@
 \end{align}
-where $S_{\mu\nu}^{(p)}\equiv\op{\mu}{\nu}_p$ flips the state of spin
-$p$ to $\ket\mu$ from $\ket\nu$; the operator $\Pi_{pq}$ permutes
-spins $p$ and $q$; and the $h_{pq}$ in the Hamiltonian are all
-negative: $h_{pq}<0$.
+where the coefficients $h_{pq}$ in the Hamiltonian are all negative,
+with $h_{pq}=h_{qp}$ and $h_{pp}=0$;
+$S_{\mu\nu}^{(p)}\equiv\op{\mu}{\nu}_p$ flips the state of spin $p$ to
+$\ket\mu$ from $\ket\nu$; and the operator $\Pi_{pq}$ permutes spins
+$p$ and $q$.  The permutation operator $\Pi_{pq}$ is a multilevel
+generalization of the SU(2)-symmetric spin interaction
+$\m S_p\c\m S_q$.
 
@@
 consists of fully symmetric states that are simultaneous $+1$
-eigenstates of all permutation operators $\Pi_{pq}$.  We will assume
-throughout these notes that the interaction Hamiltonian $H_0$ is
-gapped, i.e.~that there is a finite energy difference between $\M_0$
-and its orthogonal complement.  The energy of fully symmetric states
-$\ket\psi\in\M_0$ is
-\begin{align}
-  E_0 \equiv \sum_{p<q} h_{pq}
-  = \f12 \sum_{p,q} h_{pq}
-  = \f12 \sum_p h_p
-  = \f12 N h,
-\end{align}
-where for convenience define coefficients $h_{pq}$ for all $p,q$ (as
-opposed to only $p<q$) by enforcing $h_{pq}=h_{qp}$ and $h_{pp}=0$;
-and
-\begin{align}
-  h_p \equiv \sum_q h_{pq},
-  &&
-  h \equiv \EE_p\sp{h_p} = \f1N \sum_p h_p.
-\end{align}
-We wish to determine the effective dynamics induced on the fully
-symmetric manifold $\M_0$ by weak single- and two-body perturbations
-of the form
+eigenstates of all permutation operators $\Pi_{pq}$.  We wish to
+determine the effective dynamics induced on the fully symmetric
+manifold $\M_0$ by weak single- and two-body perturbations of the form
 \begin{align}
@@
   \V_2 \equiv \sum_{p<q} w_{pq} X_p Y_q
-  = \f12 \sum_{p,q} w_{pq} X_p Y_q
+  = \f12 \sum_{p,q} X_p w_{pq} Y_q
   = \f12 \v X \c\m w \c\v Y,
@@
 we enforce $w_{pq}=w_{qp}$ and $w_{pp}=0$; and we use a dot ($\c$) to
-denote an inner product between matrices or vectors as appropriate.
+denote an inner product between vectors and matrices as appropriate.
 
+We will assume throughout these notes that the interaction Hamiltonian
+$H_0$ is gapped, and that the interaction energy difference between
+$\M_0$ and its orthogonal complement is much larger than than any
+additional couplings that we introduce.  The energy of fully symmetric
+states $\ket\psi\in\M_0$ is
+\begin{align}
+  E_0 = \f12 \sum_{p,q} h_{pq}
+  = \f12 \sum_p h_p
+  = \f12 N h,
+\end{align}
+where for later convenience we define the net interaction energy $h_p$
+associated with spin $p$, the mean $h$ of all such energies:
+\begin{align}
+  h_p \equiv \sum_q h_{pq},
+  &&
+  h \equiv \EE_p\sp{h_p} = \f1N \sum_p h_p.
+\end{align}
 The effective Hamiltonian $H_M$ induced on the ground-state manifold
@@
 where $\P_\Delta$ is a projector onto the eigenspace of the
-interaction Hamiltonian $H_0$, with interaction energy $\Delta$
-above that of fully symmetric manifold $\M_0$.
+interaction Hamiltonian $H_0$, with interaction energy $\Delta$ above
+that of fully symmetric manifold $\M_0$; that is, the energy of states
+in the image of $\P_\Delta$ is $E_0+\Delta$.
 
@@
 \end{align}
-where $\m h$ is a matrix of all coefficients $h_{pq}$ in the
+where $\m h$ is a matrix of all coefficients $h_{pq}$ of the
 interaction Hamiltonian $H_0$ in \eqref{eq:ints}, and $\v v\oc\v h$ is
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Two-body product at first order}
+\subsection{Two-body product}
 \label{sec:PXYP}

### 2019-11-01 collective_sun/sun_perturbation.tex
more rearrangements / polishing of the text

@@
   H_0 = \sum_{p<q} h_{pq} \Pi_{pq}
-  = \f12 \sum_{p,q} \Pi_{pq},
+  = \f12 \sum_{p,q} h_{pq} \Pi_{pq},
   &&
@@
 \end{align}
-where $V_p,X_p,Y_p$ are trace-zero operators on spin $p$;
+where $V_p,X_p,Y_p$ are trace-zero operators on spin $p$, with
+e.g.~$V_p$ related to $V_q$ by a permutation of spins $p$ and $q$;
 $\v V,\v X,\v Y$ are respectively vectors of all $V_p,X_p,Y_p$; $\v v$
@@
 \begin{align}
-  E_0 = \f12 \sum_{p,q} h_{pq}
-  = \f12 \sum_p h_p
-  = \f12 N h,
-\end{align}
-where for later convenience we define the net interaction energy $h_p$
-associated with spin $p$, the mean $h$ of all such energies:
-\begin{align}
-  h_p \equiv \sum_q h_{pq},
-  &&
-  h \equiv \EE_p\sp{h_p} = \f1N \sum_p h_p.
+  E_0 = \sum_{p<q} h_{pq}.
 \end{align}
@@
   &&
-  v \equiv \EE_p\sp{v_p},
+  v \equiv \EE_p\sp{v_p} = \f1N \sum_p v_p,
   &&
@@
 where $\simeq$ denotes equality up to a restriction to the
-ground-state manifold $\M_0$.  In order to compute the second order
-effective Hamiltonian $H_1^{(2)}$, we first choose an arbitrary fully
-symmetric state $\ket\psi\in\M_0$ and expand (see Appendix
-\ref{sec:H_V1_psi}):
+ground-state manifold $\M_0$, and $\EE_p\sp{x_p}$ denotes the mean
+value of $x_p$.  In order to compute the second order effective
+Hamiltonian $H_1^{(2)}$, we first choose an arbitrary fully symmetric
+state $\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_V1_psi}):
 \begin{align}
@@
 where $\m h$ is a matrix of all coefficients $h_{pq}$ of the
-interaction Hamiltonian $H_0$ in \eqref{eq:ints}, and $\v v\oc\v h$ is
-the element-wise (Hadamard) product of $\v v$ and $\v h$, i.e.~a
-vector with components $v_kh_k$.  We thus find that the vector
-$\V_1\ket\psi$ is an eigenvector of the interaction Hamiltonian $H_0$
-if the coefficients $\v v$ satisfy
+interaction Hamiltonian $H_0$ in \eqref{eq:ints}; $\v h$ is a vector
+of the net interaction energies $h_p\equiv\sum_qh_{pq}$ associated
+with individual spins $p$; and $\v v\oc\v h$ is the element-wise
+(Hadamard) product of $\v v$ and $\v h$, i.e.~a vector with components
+$v_kh_k$.  We thus find that the vector $\V_1\ket\psi$ is an
+eigenvector of the interaction Hamiltonian $H_0$ if the coefficients
+$\v v$ satisfy
 \begin{align}
@@
 on $\v v$.  If the interaction Hamiltonian $H_0$ is translationally
-invariant, then all $h_q=h$ and $\v h\oc\v v=h\v v$, so the condition
-in \eqref{eq:cond_1} simplifies to
+invariant, then all $h_p=h\equiv\EE_q\sp{h_q}$ and
+$\v h\oc\v v=h\v v$, so the condition in \eqref{eq:cond_1} simplifies
+to
 \begin{align}
@@
   H_0 \V_1\p{\v v_\Delta} \ket\psi
-  = \sp{E_0 + \Delta} \V_1\p{\v v_\Delta} \ket\psi.
+  = \p{E_0 + \Delta} \V_1\p{\v v_\Delta} \ket\psi.
   \label{eq:H_V1_psi}
@@
 Interestingly, the energy of the (unnormalized) state
-$\V_1\p{\v v}\ket\psi$ depends only on the coefficients $\v v$, and is
-entirely independent of the state $\ket\psi$ or choice of single-spin
-trace-zero operator used to build $\V_1$ in \eqref{eq:perturbations}.
-Finding operators $\V_1\p{\v v}$ that generate eigenvectors of the
-interaction Hamiltonian $H_0$ when they act on fully symmetric states
-$\ket\psi$ thus reduces to finding eigenvectors $\v v$ of the
-interaction matrix $\m h-\diag\v h$.
+$\V_1\p{\v v_\Delta}\ket\psi$ depends only on the coefficients
+$\v v_\Delta$, and is entirely independent of the state $\ket\psi$ or
+choice of single-spin trace-zero operator $V_p$ used to build $\V_1$
+in \eqref{eq:perturbations}.  Finding operators $\V_1\p{\v v_\Delta}$
+that generate eigenvectors of the interaction Hamiltonian $H_0$ when
+they act on fully symmetric states $\ket\psi$ thus reduces to finding
+eigenvectors $\v v_\Delta$ of the interaction matrix $\m h-\diag\v h$.
 
@@
 $\V_1\p{\v v_\Delta}$ preserves the permutational symmetry of $\M_0$.
-The only way for $\V_1\p{\v v_\Delta}$ to obey permutational symmetry
-is for $\v v_\Delta$ to be a constant vector,
-i.e.~$\v v_\Delta=v_\Delta\v\1$ with $\1$ a constant vector of ones.
-Indeed, a constant vector $\v\1$ satisfies the condition in
-\eqref{eq:cond_1} with eigenvalue $0$.  All other vectors
-$\v v_\Delta$ satisfying \eqref{eq:cond_1} with $\Delta\ne0$ must be
-orthogonal to $\v\1$, and therefore mean-zero.
+Indeed, a constant vector $\v\1\equiv\p{1,1,1,\cdots}$ of ones
+satisfies the condition in \eqref{eq:cond_1} with eigenvalue $0$.  All
+other vectors $\v v_\Delta$ satisfying \eqref{eq:cond_1} with
+$\Delta\ne0$ must be orthogonal to $\v\1$, and therefore mean-zero.
 
@@
 where each operator $\V_1\p{\v v_\Delta}$ generates a state with
-interaction energy $\Delta$ above that of the fully symmetric
+definite interaction energy $\Delta$ above that of the fully symmetric
 manifold.  We can therefore simplify
@@
 \end{align}
-where the the product $\P_0 \V_1\p{\v v_\Delta}^2 \P_0$ is worked out
-in Appendix \ref{sec:PVVP}, giving us
+where the product $\P_0 \V_1\p{\v v_\Delta}^2 \P_0$ is worked out in
+Appendix \ref{sec:PVVP}, giving us
 \begin{align}
-  H_1^{(2)} \simeq \f1{N-1} \sum_{\Delta>0}
-  \f{\var\p{\v v_\Delta}}{\Delta} \p{V^2 - N W}.
+  H_1^{(2)} \simeq \f{\tilde v}{N-1} \p{V^2 - N W},
+  &&
+  \tilde v \equiv \sum_{\Delta>0} \f{\var\p{\v v_\Delta}}{\Delta}.
   \label{eq:H_1_2}
@@
   &&
-  \tilde Z \equiv \f12 \sp{X,Y}_+
+  \tilde Z \equiv \f12 \sp{X,Y}_+,
   &&
@@
 \end{align}
-where $\v\1$ is an $N$-component vector of ones,
+where $\v\1$ is an $N$-component vector of ones;
 \begin{align}
   \bar{\m h}
-  \equiv \m h - \f12\p{\v h\otimes\v\1 + \v\1\otimes\v h},
+  \equiv \m h - \f12\p{\v h\otimes\v\1 + \v\1\otimes\v h}
 \end{align}
@@
 \end{align}
-for some constant $\Delta\p{\m w}$ that may depend on $\m w$; $\v\1$
-is a constant vector of ones.  Remembering that both $\m h$ and $\m w$
-are both symmetric (under transposition) with zeros on the diagonal,
-the diagonal components of the condition in \eqref{eq:cond_2}
-automatically enforce that the last term in \eqref{eq:H_V2_psi}
-vanishes.
+for some constant $\Delta\p{\m w}$ that may depend on $\m w$.
+Remembering that the matrices $\m h$ and $\m w$ are both symmetric
+(under transposition) with zeros on the diagonal, the diagonal
+components of the condition in \eqref{eq:cond_2} automatically enforce
+that the last term ($\sim\v Z$) in \eqref{eq:H_V2_psi} vanishes.
 
-We can re-cast the condition in \eqref{eq:cond_2} as an eigenvalue
+We can recast the condition in \eqref{eq:cond_2} as an eigenvalue
 problem by writing the coefficient matrix $\m w$ as a vector in
@@
 which reduces the eigenvalue equation in \eqref{eq:cond_2_eig} to the
-$N$ smaller, uncoupled eigenvalue equations
+$N$ smaller, independent eigenvalue equations
 \begin{align}
@@
 \end{align}
-We can thus (numerically) diagonalize the blocks $\m h_\ell$
-independently, and classify their eigenvectors $\v w_\ell$ by the
-corresponding eigenvalues $\Delta\p{\m w}/2$.  ``Total'' eigenvectors
-$\v{\m w}$ of $\check{\m h}$ are then built from eigenvectors
-$\v{\m w}_\ell$ within each block $\m h_\ell$ that share the same
-eigenvalue $\Delta\p{\m w}/2$.  The task of constructing eigenvectors
+We can thus (numerically) diagonalize each of the blocks $\m h_\ell$,
+and classify their eigenvectors $\v w_\ell$ by the corresponding
+eigenvalues $\Delta\p{\m w}/2$.  ``Total'' eigenvectors $\v{\m w}$ of
+$\check{\m h}$ are then built from eigenvectors $\v w_\ell$ within
+each block $\m h_\ell$ that share the same eigenvalue
+$\Delta\p{\m w}/2$.  The task of constructing eigenvectors
 $\V_2\p{\m w}\ket\psi$ of $H_0$ and determining their energy thus
@@
 If we {\it construct} a perturbation $\V_2\p{\m w_\Delta}$ with
-coefficients $\m w_\Delta$ that satisfy \eqref{eq:cond_1} with
-eigenvalue $\Delta/2$, then
+coefficients $\m w_\Delta$ that satisfy \eqref{eq:cond_2} or
+\eqref{eq:cond_2_eig} with eigenvalue $\Delta/2$, then
 \begin{align}
   H_0 \V_2\p{\m w_\Delta} \ket\psi
-  = \sp{E_0 + \Delta} \V_2\p{\m w_\Delta} \ket\psi.
+  = \p{E_0 + \Delta} \V_2\p{\m w_\Delta} \ket\psi,
   \label{eq:H_V1_psi}
 \end{align}
+which implies that $\V_2\p{\m w_\Delta}\ket\psi$ is a state of
+definite energy that depends only on the coefficients $\m w_\Delta$.
 As with the case of single-body perturbations, we can decompose any
@@
 \begin{align}
-  \V_2\p{\m w} = \sum_\Delta \V_2\p{\m w_\Delta}.
+  \V_2\p{\m w} = \sum_\Delta \V_2\p{\m w_\Delta},
 \end{align}
-We can then simplify
+and in turn simplify
 \begin{align}
   H_2^{(2)} = - \P_0 \V_2\p{\m w} \E \V_2\p{\m w} \P_0
-  = -\sum_{\Delta>0} \f1\Delta \P_0 \V_2\p{\m w_\Delta}^2 \P_0,
+  = -\sum_{\Delta>0} \f1\Delta \P_0 \V_2\p{\m w_\Delta}^2 \P_0.
 \end{align}
-where the product $\P_0 \V_2\p{\m w_\Delta}^2 \P_0$ can be written in
-terms collective operators using techniques similar to those in
-Appendix \ref{sec:sym_prod}.
+The product $\P_0 \V_2\p{\m w_\Delta}^2 \P_0$ can be written in terms
+collective operators using techniques similar to those in Appendix
+\ref{sec:sym_prod}, which should yield a effective Hamiltonian of the
+form
+\begin{align}
+  H_2^{(2)} \simeq f\p{\m w_\Delta} \O\p{\v X,\v Y},
+\end{align}
+where $f\p{\m w_\Delta}$ is a scalar and $\O\p{\v X,\v Y}$ is a
+collective operator that is independent of $\m w_\Delta$.
 
@@
 \end{align}
-which has terms with $p,q\notin\set{k,\ell}$, one term with
-$\set{p,q}=\set{k,\ell}$, and terms with $p,q\in\set{k,\ell}$.  The
+which has one term with $\set{p,q}=\set{k,\ell}$, terms with
+$p,q\notin\set{k,\ell}$, and terms with $p,q\in\set{k,\ell}$.  The
 permutation operator $\Pi_{pq}$ acts trivially on
-$X_k Y_\ell \ket\psi$ when $p,q\notin\set{k,\ell}$ or
-$\set{p,q}=\set{k,\ell}$, leaving a sum over $p,q$ of the form
+$X_k Y_\ell \ket\psi$ when $\set{p,q}=\set{k,\ell}$ or
+$p,q\notin\set{k,\ell}$, leaving a sum over $p,q$ of the form
 \begin{align}
-  \sum_{\substack{p<q\\p,q\notin\set{k,\ell}}} h_{pq} + h_{k\ell}
+  h_{k\ell} + \sum_{\substack{p<q\\p,q\notin\set{k,\ell}}} h_{pq}
   = \sum_{p<q} h_{pq}
@@
   \P_0 \V_2 \P_0
-  = \f12 \P_0 \v X\c\m w\c\v Y\P_0
   = \f12 \sum_{p\ne q} w_{pq} \P_0 X_p Y_q \P_0
@@
 \begin{align}
+  X \equiv \sum_p X_p,
+  &&
+  Y \equiv \sum_p Y_p,
+  &&
   Z_p \equiv \f12\sp{X_p, Y_p}_+,
@@
 \begin{align}
-  \P_0 \V_2 \P_0
-  = \f12 w \P_0 \p{\tilde Z -  Z } \P_0.
+  \P_0 \V_2 \P_0 = \f12 w \P_0 \p{\tilde Z -  Z } \P_0.
 \end{align}

### 2019-11-01 collective_sun/sun_perturbation.tex
more rearranging / wording changes

@@
 \end{align}
-where the coefficients $h_{pq}$ in the Hamiltonian are all negative,
-with $h_{pq}=h_{qp}$ and $h_{pp}=0$;
+where the coefficients $h_{pq}$ in the Hamiltonian satisfy
+$h_{pq}=h_{qp}$ and $h_{pp}=0$;
 $S_{\mu\nu}^{(p)}\equiv\op{\mu}{\nu}_p$ flips the state of spin $p$ to
@@
 
-The ground-state manifold $\M_0$ of the interaction Hamiltonian $H_0$
-consists of fully symmetric states that are simultaneous $+1$
-eigenstates of all permutation operators $\Pi_{pq}$.  We wish to
-determine the effective dynamics induced on the fully symmetric
-manifold $\M_0$ by weak single- and two-body perturbations of the form
+We wish to determine the effective dynamics induced on the
+ground-state manifold $\M_0$ of $H_0$ by weak single- and two-body
+perturbations of the form
 \begin{align}
@@
 
-We will assume throughout these notes that the interaction Hamiltonian
-$H_0$ is gapped, and that the interaction energy difference between
-$\M_0$ and its orthogonal complement is much larger than than any
-additional couplings that we introduce.  The energy of fully symmetric
-states $\ket\psi\in\M_0$ is
-\begin{align}
-  E_0 = \sum_{p<q} h_{pq}.
-\end{align}
+If the coefficients $h_{pq}$ of the interaction Hamiltonian $H_0$ are
+all negative, then the ground-state manifold $\M_0$ of the interaction
+Hamiltonian $H_0$ consists of fully symmetric states that are
+simultaneous $+1$ eigenstates of all permutation operators $\Pi_{pq}$.
+We will adopt the restriction that all $h_{pq}<0$ throughout these
+notes, but note that this restriction can be relaxed to the assumption
+that the initial state of the spins is fully symmetric (e.g.~a
+spin-polarized state).  We will also assume throughout these notes
+that the fully symmetric manifold $\M_0$ is gapped away from all other
+states by an interaction energy difference that is large compared to
+any coupling between $\M_0$ and its orthogonal complement.
+
 The effective Hamiltonian $H_M$ induced on the ground-state manifold
@@
 interaction Hamiltonian $H_0$, with interaction energy $\Delta$ above
-that of fully symmetric manifold $\M_0$; that is, the energy of states
-in the image of $\P_\Delta$ is $E_0+\Delta$.
+that of fully symmetric manifold $\M_0$; that is, the interaction
+energy of states in the image of $\P_\Delta$ is $E_0+\Delta$, where
+\begin{align}
+  E_0 = \sum_{p<q} h_{pq}
+\end{align}
+is the interaction energy of fully symmetric states $\ket\psi\in\M_0$.
 
@@
 
-If the interaction Hamiltonian $H_0$ is gapped, then a vector
-$\V_1\p{\v v_\Delta}\ket\psi$ with $\Delta=0$ must lie within the
-fully symmetric manifold $\M_0$, which implies that the operator
-$\V_1\p{\v v_\Delta}$ preserves the permutational symmetry of $\M_0$.
-Indeed, a constant vector $\v\1\equiv\p{1,1,1,\cdots}$ of ones
-satisfies the condition in \eqref{eq:cond_1} with eigenvalue $0$.  All
-other vectors $\v v_\Delta$ satisfying \eqref{eq:cond_1} with
-$\Delta\ne0$ must be orthogonal to $\v\1$, and therefore mean-zero.
+If the fully symmetric manifold $\M_0$ is gapped away from all other
+states, then then a vector $\V_1\p{\v v_\Delta}\ket\psi$ with
+$\Delta=0$ must lie within the fully symmetric manifold $\M_0$, which
+implies that the operator $\V_1\p{\v v_\Delta}$ preserves the
+permutational symmetry of $\M_0$.  Indeed, a constant vector
+$\v\1\equiv\p{1,1,1,\cdots}$ of ones satisfies the condition in
+\eqref{eq:cond_1} with eigenvalue $0$.  All other vectors
+$\v v_\Delta$ satisfying \eqref{eq:cond_1} with $\Delta\ne0$ must be
+orthogonal to $\v\1$, and therefore mean-zero.
 

### 2019-11-01 collective_sun/sun_perturbation.tex
minor wording changes

@@
 \newcommand{\z}{\text{z}}
-
-\newcommand{\e}{\varepsilon}
 
@@
 We can recast the condition in \eqref{eq:cond_2} as an eigenvalue
-problem by writing the coefficient matrix $\m w$ as a vector in
-$\RR^{N^2}$,
+problem by writing the coefficient matrix $\m w$ as a vector
+$\v{\m w}$ in $\RR^N\times\RR^N$,
 \begin{align}
@@
 \end{align}
-We can thus (numerically) diagonalize each of the blocks $\m h_\ell$,
-and classify their eigenvectors $\v w_\ell$ by the corresponding
+We can (numerically) diagonalize each of the blocks $\m h_\ell$, and
+classify their eigenvectors $\v w_\ell$ by the corresponding
 eigenvalues $\Delta\p{\m w}/2$.  ``Total'' eigenvectors $\v{\m w}$ of
@@
 $\Delta\p{\m w}/2$.  The task of constructing eigenvectors
-$\V_2\p{\m w}\ket\psi$ of $H_0$ and determining their energy thus
-reduces to finding the eigenvectors and eigenvalues of $N$ matrices
-$\m h_\ell$ of size $N\times N$.
+$\V_2\p{\m w_\Delta}\ket\psi$ of $H_0$ and determining their energy
+thus reduces to finding the eigenvectors and eigenvalues of $N$
+matrices $\m h_\ell$ of size $N\times N$.
 

### 2019-11-03 collective_sun/multilevel_spin_notes.tex
simplified derivation of effective SOC hamiltonian

@@
 \renewcommand{\S}{\mathcal{S}}
+\newcommand{\V}{\mathcal{V}}
 \newcommand{\X}{\mathcal{X}}
@@
 important feature of $\M_\FS$ is its closure under the action of
-unitaries generated by collective operators of the form
-$\Q=\sum_pQ^{(p)}$, as well as linear combinations and products
-thereof.  Closure of $\M_\FS$ under collective dynamics is a
-straightforward consequence of the permutational symmetry obeyed by
-collective operators.
+collective operators of the form $\Q=\sum_pQ^{(p)}$, as well as linear
+combinations and products thereof.  Closure of $\M_\FS$ under
+collective the action of collective operators, and in particular
+collective dynamics, is a straightforward consequence of the
+permutational symmetry obeyed by collective operators.
 
@@
   &&
-  \var_q\sp{X_q} \equiv \EE_p\sp{\p{X_p-\EE_q\sp{X_q}}}
-  = \EE_p\sp{X_p^2}-\EE_q\sp{X_q}^2,
+  \var_q\sp{X_q} \equiv \EE_q\sp{\p{X_q-\EE_k\sp{X_k}}}
+  = \EE_q\sp{X_q^2}-\EE_k\sp{X_k}^2,
 \end{align}
@@
 spin-squeezing Hamiltonians derived in Ref.~\cite{he2019engineering}.
-To facilitate comparison, we note that in the case of SU(2) the drive
-operator $\D_{1,0}$ is precisely the collective spin-$z$ operator
-$S_\z$ that appears in Ref.~\cite{he2019engineering}.  Furthermore,
-drive operators $D_{LM}$ vanish when $L\ge n$, so in the case of SU(2)
-the operator $\D_{2,0}$ has no contribution to the effective
-Hamiltonians in \eqref{eq:H_eff_2_phi} and \eqref{eq:H_eff_2_phi}.
+To facilitate comparison, we note that in the case of SU(2) the
+re-scaled drive operator $\xi_1\D_{1,0}$ is precisely the collective
+spin-$z$ operator $S_\z$ that appears in
+Ref.~\cite{he2019engineering}.  Furthermore, drive operators $D_{LM}$
+vanish when $L\ge n$, so in the case of SU(2) the operator $\D_{2,0}$
+has no contribution to the effective Hamiltonians in
+\eqref{eq:H_eff_1_phi} and \eqref{eq:H_eff_2_phi}.
 
@@
 symmetric manifold $\M_\FS$ by the single-particle SOC Hamiltonian
-$H_{\t{lat}}^{(\phi)}$ in \eqref{eq:H_lat_SOC_B} in the weak SOC limit
-$\abs{\phi}\ll1$.  For brevity, we work in units with the lattice
-spacing $a=1$ throughout this section.  Perturbations $V$ to the
-interaction Hamiltonian $H_{\t{int}}$ induce the following effective
-Hamiltonians on the fully symmetric manifold
-$\M_\FS$\cite{bravyi2011schrieffer, perlin2019effective}:
-\begin{align}
-  H_{\t{eff}}^{(1)} = \P_0 V \P_0,
-  &&
-  H_{\t{eff}}^{(2)} = -\P_0 V \E V \P_0,
-  &&
-  \E \equiv \sum_{k>0} \f{\P_k}{\Delta_k},
-  \label{eq:pert}
-\end{align}
-where $\P_k$ is a projector onto the $k$-th excited eigenspace of the
-interaction Hamiltonian $H_{\t{int}}$, with interaction energy
-$\Delta_k$ above that of fully symmetric states
-$\ket{\psi_\FS}\in\M_\FS$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Effective SOC at first order}
-\label{sec:H_eff_1}
-
-The single-particle SOC Hamiltonian is
 \begin{align}
@@
   \end{cases},
-  \label{eq:H_lat_SOC_A}
 \end{align}
@@
 $A_L^{(\phi)} \sim \phi^L + O\p{\phi^{L+2}}$ when $\abs{\phi}\ll1$.
-When the SOC angle $\phi$ is sufficiently small in magnitude, the
-effective Hamiltonian induced on the fully symmetric manifold $\M_\FS$
-at first order in the perturbation $H_{\t{lat}}^{(\phi)}$ is
+For brevity, we work in units with the lattice spacing $a=1$ and
+implicitly assume the limit $\abs{\phi}\ll1$ throughout this section.
+
+In order to simplify the calculation of effective Hamiltonians, we
+write the SOC Hamiltonian in the form
+\begin{align}
+  H_{\t{lat}}^{(\phi)} = \sum_{q,L} v_L^{(q)} D_{L,0}^{(q)},
+  &&
+  v_L^{(q)} \equiv - t w_L\p{q} A_L^{(\phi)},
+\end{align}
+and decompose it into its uniform and inhomogeneous parts as
+\begin{align}
+  H_{\t{lat}}^{(\phi)}
+  = \sum_L v_L \D_{L,0} + \sum_{q,L} z_L^{(q)} D_{L,0}^{(q)},
+\end{align}
+where
+\begin{align}
+  v_L \equiv \EE_q\sp{v_L^{(q)}},
+  &&
+  z_L^{(q)} \equiv v_L^{(q)} - v_L,
+  &&
+  \EE_q\sp{X_q} \equiv \f1N \sum_q X_q.
+\end{align}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Effective SOC at first order}
+\label{sec:H_eff_1}
+
+The effective Hamiltonian in induced on the fully symmetric manifold
+$\M_\FS$ at first order is
 \begin{align}
@@
   = \P_0 H_{\t{lat}}^{(\phi)} \P_0
-  = -t \sum_{q,L} w_L\p{q} A_L^{(\phi)} \P_0 D_{L,0}^{(q)} \P_0,
-\end{align}
-where we can freely change the index of each single-mode drive
-operator $D_{L,0}^{(q)}$ within the fully symmetric manifold to
-re-write
-\begin{align}
-  \P_0 D_{L,0}^{(q)} \P_0 = \f1N \sum_p \P_0 D_{L,0}^{(p)} \P_0
-  = \f1N \P_0 \D_{L,0} \P_0,
+  = \sum_L v_L \P_0 \D_{L,0} \P_0
+  + \sum_{q,L} z_L^{(q)} \P_0 D_{L,0}^{(q)} \P_0,
+\end{align}
+where $\P_0$ is a projector onto $\M_\FS$.  Permutational symmetry
+within the fully symmetric manifold $\M_\FS$ allow us to freely change
+the quasi-momentum index on $\P_0D_{L,0}^{(q)}\P_0$, and the
+coefficients $z_L^{(q)}$ are mean-zero by construction, so
+\begin{align}
+  \sum_{q,L} z_L^{(q)} \P_0 D_{L,0}^{(q)} \P_0
+  = \sum_q z_L^{(q)} \P_0 D_{L,0}^{(0)} \P_0
+  = 0,
 \end{align}
@@
   H_{\t{eff}}^{(1)}
+  = \sum_L v_L \D_{L,0}
   = -t \sum_L \EE_q\sp{w_L\p{q}} A_L^{(\phi)} \D_{L,0},
-  &&
-  \EE_q\sp{X_q} \equiv \f1N \sum_q X_q.
   \label{eq:H_eff_1}
 \end{align}
-Here $\EE_q\sp{X_q}$ is an average of $X_q$ over all occupied
-quasi-momenta $q$.  Note that we have dropped the explicit projection
-operators $\P_0$ in \eqref{eq:H_eff_1} with the understanding that we
-are implicitly restricting all effective dynamics to the fully
+where we implicitly restrict $H_{\t{eff}}^{(1)}$ to the fully
 symmetric manifold $\M_\FS$.
@@
 
-In order to expand the second-order effective Hamiltonian
-$H_{\t{eff}}^{(2)}$ induced on the fully symmetric manifold $\M_\FS$
-by the SOC Hamiltonian $H_{\t{lat}}^{(\phi)}$ in
-\eqref{eq:H_lat_SOC_A} when the SOC angle $\abs{\phi}\ll1$, we note
-that $w_L\p{q}$ in \eqref{eq:H_lat_SOC_A} takes the form
-\begin{align}
-  w_L\p{q} = \f12 \sum_{s\in\set{\pm1}} \chi_{Ls} e^{isq},
-  \label{eq:w_chi}
-\end{align}
-for some appropriately defined coefficients $\chi_{Ls}$, which implies
-that
-\begin{align}
-  H_{\t{lat}}^{(\phi)}
-  = -\f{t}{2} \sum_{\substack{L\\s\in\set{\pm1}}}
-  \chi_{Ls} A_L^{(\phi)} \Z_{Ls},
-  &&
-  \Z_{Ks} \equiv \sum_q e^{isq} D_{L,0}^{(q)},
-\end{align}
-and in turn
+The effective Hamiltonian in induced on the fully symmetric manifold
+$\M_\FS$ at second order is\cite{bravyi2011schrieffer,
+  perlin2019effective}
 \begin{align}
   H_{\t{eff}}^{(2)}
-  = -\f{t^2}{4} \sum_{\substack{J,K\\r,s\in\set{\pm1}}}
-  \chi_{Jr} \chi_{Ks} A_J^{(\phi)} A_K^{(\phi)}
-  \P_0 \Z_{Jr} \E \Z_{Ks} \P_0,
+  = - \P_0 H_{\t{lat}}^{(\phi)} \E H_{\t{lat}}^{(\phi)} \P_0
+  &&
+  \E \equiv \sum_{\Delta>0} \f{\P_\Delta}{\Delta},
+\end{align}
+where $\P_\Delta$ is a projector onto the eigenspace of the
+interaction Hamiltonian $H_0$, with interaction energy $\Delta$ above
+that of fully symmetric manifold $\M_\FS$.  The uniform part of the
+SOC Hamiltonian $H_{\t{lat}}^{(\phi)}$ is permutationally symmetric,
+and therefore does not couple the fully symmetric manifold $\M_\FS$ to
+its orthogonal complement.  As a consequence, the uniform part of
+$H_{\t{lat}}^{(\phi)}$ is annihilated between the projector $\P_0$
+onto $\M_\FS$ and the projectors in $\E$ onto the orthogonal
+complement of $\M_\FS$.  We can therefore keep only the inhomogeneous
+parts of $H_{\t{lat}}^{(\phi)}$ and expand
+\begin{align}
+  H_{\t{eff}}^{(2)}
+  = -\sum_{J,K} \P_0 \Z_J \E \Z_K \P_0,
+  &&
+  \Z_L \equiv \sum_{q,L} z_L^{(q)} D_{L,0},
   \label{eq:H_eff_2_start}
 \end{align}
-where $\P_0$ is a projector onto the fully symmetric manifold
-$\M_\FS$, and $\E$ is a sum over projectors onto excited-state
-manifolds with corresponding energetic suppression factors, defined in
-\eqref{eq:pert}.  The remainder of this section is devoted to the task
-of simplifying \eqref{eq:H_eff_2_start} to arrive at our final result
-in \eqref{eq:H_eff_2}, expressed in purely in terms of collective
-drive operators that are implicitly restricted to the fully symmetric
-manifold $\M_\FS$.  We will break up our task into three steps:
+We now divide the task of further simplifying $H_{\t{eff}}^{(2)}$ into
+two steps:
 \begin{enumerate*}
-\item eliminating the excited-state projectors in $\E$,
-\item evaluating the product of operators in \eqref{eq:H_eff_2_start},
-  and
-\item evaluating the sum over $r,s\in\set{\pm1}$.
+\item eliminating the excited-state projectors in $\E$, and
+\item projecting the remaining product of operators in
+  \eqref{eq:H_eff_2_start} onto the fully symmetric manifold.
 \end{enumerate*}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Mean-subtracted spin-wave operators and the elimination
-  of excited-state projectors}
-
-In order to simplify the product of operators in
-\eqref{eq:H_eff_2_start}, we note that sandwiching the ``spin-wave''
-operators $\Z_{Jr},\Z_{Ks}$ by the ground-state projector $\P_0$ and
-the excited-state projectors in $\E$, picks off the components of
-these spin-wave operators that are off-diagonal with respect to the
-symmetric manifold $\M_\FS$ and its orthogonal complement.  We
-therefore define the mean-subtracted spin-wave operators
-\begin{align}
-  \tilde\Z_{Ks}
-  \equiv \sum_q \p{e^{isq} - \EE_k\sp{e^{isk}}} D_{K,0}^{(q)},
-  &&
-  \EE_k\sp{X_k} \equiv \f1N \sum_k X_k,
-\end{align}
-for which the difference $\tilde\Z_{Ks}-\Z_{Ks}\propto\D_{K,0}$ is a
-collective operator that only couples fully symmetric states
-$\ket{\psi_\FS}\in\M_\FS$ to fully symmetric states.  In particular,
-\begin{align}
-  \P_0 \tilde\Z_{Jr} \E \tilde\Z_{Ks} \P_0
-  = \P_0 \Z_{Jr} \E \Z_{Ks} \P_0,
-\end{align}
-which allows us to freely take
-$\Z_{Jr},\Z_{Ks}\to\tilde\Z_{Jr},\tilde\Z_{Ks}$ with no consequence
-for the effective Hamiltonian $H_{\t{eff}}^{(2)}$ in
-\eqref{eq:H_eff_2_start}.  Mean-subtracted spin-wave operators
-$\tilde\Z_{Ks}$ have the convenient feature that their action on any
-fully symmetric state $\ket{\psi_{\FS}}\in\M_\FS$ yields an
-eigenvector of the interaction Hamiltonian
-$H_{\t{int}}=-u\sum_{p<q}\Pi_{pq}$.  To substantiate this claim, we
-expand
-\begin{align}
-  \sum_{p<q} \Pi_{pq} \tilde\Z_{Ks} \ket{\psi_\FS}
+\subsubsection{Eliminating excited-state projectors}
+
+In order to eliminate the excited-state projectors in $\E$, we observe
+that the action of $\Z_L$ on any fully symmetric state
+$\ket{\psi_{\FS}}\in\M_\FS$ yields an eigenvector of the interaction
+Hamiltonian $H_{\t{int}}=-u\sum_{p<q}\Pi_{pq}$.  To substantiate this
+claim, we expand
+\begin{align}
+  \sum_{p<q} \Pi_{pq} \Z_L \ket{\psi_\FS}
   = \sum_{\substack{p<q\\k}}
-  \p{e^{isk} - \EE_\ell\sp{e^{is\ell}}}
-  \Pi_{pq} D_{K,0}^{(k)} \ket{\psi_\FS}.
-  \label{eq:Z_eig_start}
+  z_L^{(k)} \Pi_{pq} D_{L,0}^{(k)} \ket{\psi_\FS}
 \end{align}
 If $k\notin\set{p,q}$, then the permutation operator $\Pi_{pq}$
-commutes with $D_{K,0}^{(k)}$, and in turn acts trivially on the
+commutes with $D_{L,0}^{(k)}$, and in turn acts trivially on the
 permutationally-symmetric state $\ket{\psi_\FS}$, giving us
@@
   \sum_{\substack{p<q\\k\notin\set{p,q}}}
-  \p{e^{isk} - \EE_\ell\sp{e^{is\ell}}}
-  \Pi_{pq} D_{K,0}^{(k)} \ket{\psi_\FS}
-  = { N-1 \choose 2 } \tilde\Z_{Ks} \ket{\psi_\FS}.
+  z_L^{(k)} \Pi_{pq} D_{L,0}^{(k)} \ket{\psi_\FS}
+  = \sum_k \sum_{\substack{p<q\\p,q\ne k}}
+  z_L^{(k)} D_{L,0}^{(k)} \ket{\psi_\FS}
+  = { N-1 \choose 2 } \Z_L \ket{\psi_\FS}.
   \label{eq:Z_eig_notin}
@@
 If $k\in\set{p,q}$, then we can resolve the identity operator
-$\1=\Pi_{pq}^\dag \Pi_{pq}$ to expand
+$\1=\Pi_{pq}\Pi_{pq}$ to expand
 \begin{multline}
   \sum_{\substack{p<q\\k\in\set{p,q}}}
-  \p{e^{isk} - \EE_\ell\sp{e^{is\ell}}}
-  \Pi_{pq} D_{K,0}^{(k)} \ket{\psi_\FS}
+  z_L^{(k)} \Pi_{pq} D_{L,0}^{(k)} \ket{\psi_\FS}
   = \f12 \sum_{\substack{p\ne q\\k\in\set{p,q}}}
-  \p{e^{isk} - \EE_\ell\sp{e^{is\ell}}}
-  \Pi_{pq} D_{K,0}^{(k)} \Pi_{pq}^\dag \Pi_{pq} \ket{\psi_\FS} \\
-  = \sum_q \sum_{k\ne q} \p{e^{isk} - \EE_\ell\sp{e^{is\ell}}}
-  D_{K,0}^{(q)} \ket{\psi_\FS},
+  z_L^{(k)} \Pi_{pq} D_{L,0}^{(k)} \Pi_{pq} \Pi_{pq} \ket{\psi_\FS}
+  = \sum_q \sum_{k\ne q} z_L^{(k)} D_{L,0}^{(q)} \ket{\psi_\FS},
 \end{multline}
-where the sum over $k\ne q$ is equal to a sum over all $k$ minus a
-term with $k=q$, i.e.
-\begin{align}
-  \sum_{k\ne q} \p{e^{isk} - \EE_\ell\sp{e^{is\ell}}}
-  = \sum_k \p{e^{isk} - \EE_\ell\sp{e^{is\ell}}}
-  - \p{e^{isq} - \EE_\ell\sp{e^{is\ell}}}
-  = - \p{e^{isq} - \EE_\ell\sp{e^{is\ell}}},
+where, using the fact that the coefficients $z_L^{(k)}$ are mean-zero,
+\begin{align}
+  \sum_{k\ne q} z_L^{(k)}
+  = \sum_k z_L^{(k)} - z_L^{(q)}
+  = - z_L^{(q)},
 \end{align}
@@
   \sum_{\substack{p<q\\k\in\set{p,q}}}
-  \p{e^{isk} - \EE_\ell\sp{e^{is\ell}}}
-  \Pi_{pq} D_{K,0}^{(k)} \ket{\psi_\FS}
-  = - \tilde\Z_{Ks} \ket{\psi_\FS}.
-  \label{eq:Z_eig_in}
-\end{align}
-Substituting the results in \eqref{eq:Z_eig_in} and
-\eqref{eq:Z_eig_notin} into \eqref{eq:Z_eig_start} gives us
-\begin{align}
-  \sum_{p<q} \Pi_{pq} \tilde\Z_{Ks} \ket{\psi_\FS}
-  = \sp{{ N-2 \choose 2 } - 1} \tilde\Z_{Ks} \ket{\psi_\FS}.
+  z_L^{(k)} \Pi_{pq} D_{L,0}^{(k)} \ket{\psi_\FS}
+  = - \sum_q z_L^{(q)} D_{L,0}^{(q)} \ket{\psi_\FS}
+  = - \Z_L \ket{\psi_\FS}.
+\end{align}
+In total,
+\begin{align}
+  \sum_{p<q} \Pi_{pq} \Z_L \ket{\psi_\FS}
+  = \sp{{N-1\choose 2}-1} \Z_L \ket{\psi_\FS}.
 \end{align}
 The image of any fully symmetric state $\ket{\psi_\FS}\in\M_\FS$ under
-the action of the mean-subtracted spin-wave operator $\tilde\Z_{Ks}$
-is thus proportional to an eigenstate of $H_{\t{int}}$ with
-interaction energy $-u\sp{{ N-1 \choose 2 } - 1}$, which is $Nu$ above
-the interaction energy of $-u{N\choose2}$ of states in the fully
-symmetric manifold, so
-\begin{align}
-  \E \tilde\Z_{Ks} \P_0
-  =  \f1{Nu} \tilde\Z_{Ks} \P_0
-  = \f1{fU} \tilde\Z_{Ks} \P_0,
+the action of $\Z_L$ is thus proportional to an eigenstate of
+$H_{\t{int}}$ with interaction energy $-u\sp{{ N-1 \choose 2 } - 1}$,
+which is $Nu$ above the interaction energy $-u{N\choose2}$ of states
+in the fully symmetric manifold, so
+\begin{align}
+  \E \Z_L \P_0
+  = \sum_{\Delta>0} \f1{\Delta} \P_\Delta \Z_L \P_0
+  =  \f1{Nu} \Z_L \P_0
+  = \f1{fU} \Z_L \P_0,
   \label{eq:proj_elim}
 \end{align}
-where $f$ is the filling fraction of lattice sites.  We have thus far
-simplified the effective Hamiltonian in \eqref{eq:H_eff_2_start} to
+where $f$ is the filling fraction of lattice sites and $U$ is the
+two-body on-site interaction energy.  Our effective Hamiltonian thus
+far is
 \begin{align}
   H_{\t{eff}}^{(2)}
-  = -\f{t^2}{4fU} \sum_{\substack{J,K\\r,s\in\set{\pm1}}}
-  \chi_{Jr} \chi_{Ks} A_J^{(\phi)} A_K^{(\phi)}
-  \P_0 \tilde\Z_{Jr} \tilde\Z_{Ks} \P_0,
+  = - \f1{fU} \sum_{J,K} \P_0 \Z_J \Z_K \P_0.
   \label{eq:H_eff_2_mid}
 \end{align}
-leaving us the task of evaluating the product
-$\P_0 \tilde\Z_{Jr} \tilde\Z_{Ks} \P_0$, as well as the sum over
-$r,s$.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{The spin-wave operator product}
-
-To evaluate the product $\P_0 \tilde\Z_{Jr} \tilde\Z_{Ks} \P_0$ of
-mean-subtracted spin-wave operators projected by $\P_0$ onto the fully
-symmetric manifold $\M_\FS$, we expand
-\begin{align}
-  \P_0 \tilde\Z_{Jr} \tilde\Z_{Ks} \P_0
-  = \sum_{p,q} \p{e^{irp}-\EE_k\sp{e^{irk}}}
-  \p{e^{isq}-\EE_\ell\sp{e^{is\ell}}}
-  \P_0 D_{J,0}^{(p)} D_{K,0}^{(q)} \P_0.
-  \label{eq:ZZ_prod_start}
+\subsubsection{Projecting onto the fully symmetric manifold}
+
+We now need to evaluate the product
+\begin{align}
+  \P_0 \Z_J \Z_K \P_0
+  = \sum_{p,q} z_J^{(p)} z_K^{(q)} \P_0 D_{J,0}^{(p)} D_{K,0}^{(q)} \P_0.
+  \label{eq:PZZP_start}
 \end{align}
@@
 \begin{align}
-  \sum_p \p{e^{irp}-\EE_k\sp{e^{irk}}}
-  \p{e^{isp}-\EE_\ell\sp{e^{is\ell}}}
-  \P_0 D_{J,0}^{(p)} D_{K,0}^{(p)} \P_0
-  = N \cov_{p,q}\sp{e^{irp},e^{isq}}
+  \sum_q z_J^{(q)} z_K^{(q)}
+  \P_0 D_{J,0}^{(q)} D_{K,0}^{(q)} \P_0
+  = N \cov_q\sp{v_J^{(q)},v_K^{(q)}}
   \P_0 D_{J,0}^{(0)} D_{K,0}^{(0)} \P_0,
+  \label{eq:PZZP_eq}
 \end{align}
@@
 \begin{align}
-  \cov_{p,q}\sp{X_p,Y_q}
-  \equiv \EE_p\sp{\p{X_p-\EE_k\sp{X_k}}\p{Y_p-\EE_\ell\sp{Y_\ell}}}
-  = \EE_k\sp{X_k Y_k} - \EE_p\sp{X_p} \EE_q\sp{X_q}
+  \cov_q\sp{X_q,Y_q}
+  \equiv \EE_q\sp{\p{X_q-\EE_k\sp{X_k}}\p{Y_q-\EE_\ell\sp{Y_\ell}}}
+  = \EE_q\sp{X_q Y_q} - \EE_k\sp{X_k} \EE_\ell\sp{Y_\ell}
   \label{eq:cov}
 \end{align}
-is the covariance between $X_p$ and $Y_q$, evaluated over all occupied
-quasi-momenta $p,q$.  Similarly, the $p\ne q$ terms in
-\eqref{eq:ZZ_prod_start} simplify to
-\begin{multline}
-  \sum_{p\ne q} \p{e^{irp}-\EE_k\sp{e^{irk}}}
-  \p{e^{isq}-\EE_\ell\sp{e^{is\ell}}}
-  \P_0 D_{J,0}^{(p)} D_{K,0}^{(q)} \P_0 \\
-  = \sum_p \p{e^{irp}-\EE_k\sp{e^{irk}}}
-  \sum_{q\ne p} \p{e^{isq}-\EE_\ell\sp{e^{is\ell}}}
+is the covariance between $X_q$ and $Y_q$, evaluated over all occupied
+quasi-momenta $q$.  The terms with $p\ne q$ in \eqref{eq:PZZP_start},
+meanwhile, are
+\begin{align}
+  \sum_{p\ne q} z_J^{(p)} z_K^{(q)}
+  \P_0 D_{J,0}^{(p)} D_{K,0}^{(q)} \P_0
+  = \sum_{p\ne q} z_J^{(p)} z_K^{(q)}
   \P_0 D_{J,0}^{(0)} D_{K,0}^{(1)} \P_0,
-\end{multline}
-where the sum over $q\ne p$ is equal to a sum over all $q$ minus a
-term with $q=p$, i.e.
-\begin{align}
-  \sum_{q\ne p} \p{e^{isq}-\EE_\ell\sp{e^{is\ell}}}
-  = \sum_q \p{e^{isq}-\EE_\ell\sp{e^{is\ell}}}
-  - \p{e^{isp}-\EE_\ell\sp{e^{is\ell}}}
-  = - \p{e^{isp}-\EE_\ell\sp{e^{is\ell}}},
-\end{align}
-which implies that the overall sum over $p,q$ is
-\begin{align}
-  -\sum_p \p{e^{irp}-\EE_k\sp{e^{irk}}}
-  \p{e^{isp}-\EE_\ell\sp{e^{is\ell}}}
-  = - N \cov_{p,q}\sp{e^{irp},e^{isq}}.
-\end{align}
-Thus far, we have that
-\begin{align}
-  \P_0 \tilde\Z_{Jr} \tilde\Z_{Ks} \P_0
-  = N \cov_{p,q}\sp{e^{irp},e^{isq}}
-  \p{\P_0 D_{J,0}^{(0)} D_{K,0}^{(0)} \P_0
-    - \P_0 D_{J,0}^{(0)} D_{K,0}^{(1)} \P_0}.
-  \label{eq:ZZ_prod_mid}
-\end{align}
-In order to write this result in terms of collective operators, we
+\end{align}
+where the remaining sum over coefficients is
+\begin{align}
+  \sum_{p\ne q} z_J^{(p)} z_K^{(q)}
+  = \sum_{p,q} z_J^{(p)} z_K^{(q)} - \sum_q z_J^{(q)} z_K^{(q)}
+  = - N \cov_q\sp{v_J^{(q)},v_K^{(q)}},
+\end{align}
+so
+\begin{align}
+  \sum_{p\ne q} z_J^{(p)} z_K^{(q)}
+  \P_0 D_{J,0}^{(p)} D_{K,0}^{(q)} \P_0
+  = - N \cov_q\sp{v_J^{(q)},v_K^{(q)}}
+  \P_0 D_{J,0}^{(0)} D_{K,0}^{(1)} \P_0.
+  \label{eq:PZZP_neq}
+\end{align}
+Combining \eqref{eq:PZZP_eq} and \eqref{eq:PZZP_neq} gives us
+\begin{align}
+  \P_0 \Z_J \Z_K \P_0
+  = N \cov_q\sp{v_J^{(q)},v_K^{(q)}}
+  \P_0\p{D_{J,0}^{(0)} D_{K,0}^{(0)}
+    - D_{J,0}^{(0)} D_{K,0}^{(1)}} \P_0.
+\end{align}
+In order to express this result in terms of collective operators, we
 expand
-\begin{multline}
+\begin{align}
   \P_0 \D_{J,0} \D_{K,0} \P_0
-  = \sum_p \P_0 D_{J,0}^{(p)} D_{K,0}^{(p)} \P_0
-  + \sum_{p\ne q} \P_0 D_{J,0}^{(p)} D_{K,0}^{(q)} \P_0 \\
   = N \P_0 D_{J,0}^{(0)} D_{K,0}^{(0)} \P_0
-  + N \p{N-1} \P_0 D_{J,0}^{(0)} D_{K,0}^{(1)} \P_0,
-\end{multline}
+  + N\p{N-1} \P_0 D_{J,0}^{(0)} D_{K,0}^{(1)} \P_0,
+\end{align}
 which implies that
@@
   - N \P_0 D_{J,0}^{(0)} D_{K,0}^{(1)} \P_0
-  = \f{N}{N-1} \P_0 D_{J,0}^{(0)} D_{K,0}^{(0)} \P_0
-  - \f1{N-1} \P_0 \D_{J,0} \D_{K,0} \P_0.
-\end{align}
-The last ingredient we need to re-write \eqref{eq:ZZ_prod_mid} in
-terms of collective operators is
+  = \f1{N-1} \P_0 \p{N D_{J,0}^{(0)} D_{K,0}^{(0)}
+    - \D_{J,0} \D_{K,0}} \P_0
+\end{align}
+and in turn
+\begin{align}
+  \P_0 \Z_J \Z_K \P_0
+  = \f1{N-1} \cov_q\sp{v_J^{(q)},v_K^{(q)}}
+  \P_0 \p{N^2 D_{J,0}^{(0)} D_{K,0}^{(0)} - \D_{J,0} \D_{K,0}} \P_0.
+\end{align}
+Finally, we can simplify
 \begin{align}
   N \P_0 D_{J,0}^{(0)} D_{K,0}^{(0)} \P_0
-  = N \sum_L f_{J,0;K,0}^{L,0} \P_0 D_{L,0}^{(0)} \P_0
-  = \sum_L f_{J,0;K,0}^{L,0} \P_0 \D_{L,0} \P_0,
-\end{align}
-where $f_{J,0;K,0}^{L,0}$ is a structure constant of the transition
-operator algebra, defined in \eqref{eq:trans_prod} of Appendix
-\ref{sec:trans_prod}.  Altogether, we then have that
-\begin{align}
-  N\p{\P_0 D_{J,0}^{(0)} D_{K,0}^{(0)} \P_0
-    - \P_0 D_{J,0}^{(0)} D_{K,0}^{(1)} \P_0}
-  = \f{N}{N-1} \sum_L f_{J,0;K,0}^{L,0} \P_0 \D_{L,0} \P_0
-  - \f1{N-1} \P_0 \D_{J,0} \D_{K,0} \P_0,
-  \label{eq:eff_sym_drive}
-\end{align}
-which, substituted into \eqref{eq:ZZ_prod_mid}, implies that
-\begin{align}
-  \P_0 \tilde\Z_{Jr} \tilde\Z_{Ks} \P_0
-  = -\f{\cov_{p,q}\sp{e^{irp},e^{isq}}}{N-1}
-  \p{\P_0 \D_{J,0} \D_{K,0} \P_0
-    - N \sum_L f_{J,0;K,0}^{L,0} \P_0 \D_{L,0} \P_0}.
-  \label{eq:ZZ_prod}
-\end{align}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Putting it all together}
-
-Having evaluated the product $\P_0 \tilde\Z_{Jr} \tilde\Z_{Ks} \P_0$
-of spin-wave operators in \eqref{eq:ZZ_prod}, the last ingredient
-needed to simplify the effective Hamiltonian $H_{\t{eff}}^{(2)}$ in
-\eqref{eq:H_eff_2_mid} is an evaluation of the sum over
-$r,s\in\set{\pm1}$.  We can use the implicit definition of the
-coefficients $\chi_{Ls}$ in \eqref{eq:w_chi} together with the
-bilinearity of the covariance, defined in \eqref{eq:cov}, to simplify
-\begin{align}
-  \f14 \sum_{r,s\in\set{\pm1}} \chi_{Jr} \chi_{Ks}
-  \cov_{p,q}\sp{e^{irp},e^{isq}}
-  = \cov_{p,q}\sp{w_J\p{p},w_K\p{q}}.
-  \label{eq:sign_sum}
-\end{align}
-Substituting the operator product in \eqref{eq:ZZ_prod} into the
-effective Hamiltonian $H_{\t{eff}}^{(2)}$ in \eqref{eq:H_eff_2_mid}
-and evaluating the sum over $r,s\in\set{\pm1}$ according to
-\eqref{eq:sign_sum} gives us
+  = N \sum_L f_{JKL} \P_0 D_{L,0}^{(0)} \P_0
+  = \sum_L f_{JKL} \P_0 \D_{L,0} \P_0,
+  &&
+  f_{JKL} \equiv f_{J,0;K,0}^{L,0},
+\end{align}
+where $f_{\ell_1 m_1;\ell_2 m_2}^{LM}$ is a structure constant of the
+transition operator algebra, defined in \eqref{eq:trans_prod}.  In
+total,
+\begin{align}
+  \P_0 \Z_J \Z_K \P_0
+  = \f1{N-1} \cov_q\sp{v_J^{(q)},v_K^{(q)}}
+  \P_0 \p{N \sum_L f_{JKL} \D_{L,0} - \D_{J,0} \D_{K,0}} \P_0,
+\end{align}
+which implies that the effective Hamiltonian $H_{\t{eff}}^{(2)}$ in
+\eqref{eq:H_eff_2_mid} is
 \begin{align}
   H_{\t{eff}}^{(2)}
-  = \f{t^2}{\p{N-1}fU} \sum_{J,K} \cov_{p,q}\sp{w_J\p{pa},w_K\p{qa}}
-  A_J^{(\phi)} A_K^{(\phi)}
-  \p{\D_{J,0} \D_{K,0} - N \sum_L f_{J,0;K,0}^{L,0} \D_{L,0}},
+  = \f{t^2}{fU\p{N-1}} \sum_{J,K}
+  \cov_q\sp{w_J\p{qa},w_K\p{qa}} A_J^{(\phi)} A_K^{(\phi)}
+  \p{\D_{J,0} \D_{K,0} - N \sum_L f_{JKL} \D_{L,0}},
   \label{eq:H_eff_2}
@@
 effective dynamics to the fully symmetric manifold $\M_\FS$.  Note
-that as the spatial filling fraction $f\to1$, such that sums over
-occupied quasi-momenta $q$ become sums over nearly all angles
-$q\in\ZZ_N\times2\pi/N$, when both $J,K$ are even or odd
-(i.e.~$J=K~\t{mod}~2$) the covariance in \eqref{eq:H_eff_2} becomes
-$\cov_{p,q}\sp{\sin\p{p},\sin\p{p}}\to1/2$ or
-$\cov_{p,q}\sp{\cos\p{p},\cos\p{p}}\to1/2$, whereas when one of $J,K$
-is even and the other is odd (i.e.~$J\ne K~\t{mod}~2$), this
-covariance becomes $\cov_{p,q}\sp{\cos\p{p},\sin\p{q}}\to0$.  In the
-case of SU(2), the effective Hamiltonian $H_{\t{eff}}^{(2)}$ in
-\eqref{eq:H_eff_2} reduces to the one-axis twisting Hamiltonian
-derived in Ref.~\cite{he2019engineering}.
+that as the spatial filling fraction $f\to1$, the sums over occupied
+quasi-momenta $q$ become sums over nearly all angles
+$q\in\ZZ_N\times2\pi/N$.  In this case, when both $J,K$ are even or
+odd (i.e.~$J=K~\t{mod}~2$) the covariance in \eqref{eq:H_eff_2}
+becomes $\cov_p\sp{\sin\p{p},\sin\p{p}}\to1/2$ or
+$\cov_p\sp{\cos\p{p},\cos\p{p}}\to1/2$, whereas when one of $J,K$ is
+even and the other is odd (i.e.~$J\ne K~\t{mod}~2$), this covariance
+becomes $\cov_p\sp{\cos\p{p},\sin\p{p}}\to0$.  In the case of SU(2),
+the effective Hamiltonian $H_{\t{eff}}^{(2)}$ in \eqref{eq:H_eff_2}
+reduces to the one-axis twisting Hamiltonian derived in
+Ref.~\cite{he2019engineering}.
 
@@
 \begin{align}
-  \var_q\sp{X_q} \equiv \cov_{p,q}\sp{X_p,X_q}
+  \var_q\sp{X_q} \equiv \cov_p\sp{X_p,X_p}
   = \EE_p\sp{\p{X_p-\EE_q\sp{X_q}}^2}
@@
 have neglected scalar terms $\sim\D_{0,0}\propto\1$ that have no
-physical consequence.  Note that in the case of SU(2), the drive
-operator $\D_{1,0}$ is precisely the collective spin-$z$ operator
-$S_\z$ that appears in Ref.~\cite{he2019engineering}.  Furthermore,
-drive operators $D_{LM}$ vanish when $L\ge n$, so in the case of SU(2)
-the operator $\D_{2,0}$ has no contribution to the effective
-Hamiltonians in \eqref{eq:H_eff_2_phi_apndx} and
+physical consequence.  Note that in the case of SU(2), the re-scaled
+drive operator $\xi_1\D_{1,0}$ is precisely the collective spin-$z$
+operator $S_\z$ that appears in Ref.~\cite{he2019engineering}.
+Furthermore, drive operators $D_{LM}$ vanish when $L\ge n$, so in the
+case of SU(2) the operator $\D_{2,0}$ has no contribution to the
+effective Hamiltonians in \eqref{eq:H_eff_2_phi_apndx} and
 \eqref{eq:H_eff_2_phi_apndx}.

### 2019-11-05 collective_sun/sun_perturbation.tex
fixed minor error in definition of two-body perturbation
minor error fixes and notation changes

@@
   &&
-  \V_2 \equiv \sum_{p<q} w_{pq} X_p Y_q
-  = \f12 \sum_{p,q} X_p w_{pq} Y_q
+  \V_2 \equiv \f12 \sum_{p,q} w_{pq} X_p Y_q
   = \f12 \v X \c\m w \c\v Y,
@@
 \begin{align}
-  H_1^{(1)} =  \sum_p v_p \P_0 V_0 \P_0 \simeq v V,
+  H_1^{(1)} =  \sum_p v_p \P_0 V_0 \P_0 \simeq \bar v V,
   &&
-  v \equiv \EE_p\sp{v_p} = \f1N \sum_p v_p,
+  \bar v \equiv \EE_p\sp{v_p} = \f1N \sum_p v_p,
   &&
@@
   = \f12 \P_0 \v X\c\m w\c\v Y\P_0
-  \simeq \f12 w \p{\tilde Z - Z},
+  \simeq \f12 \bar w \p{\tilde Z - Z},
 \end{align}
@@
 \begin{align}
-  w \equiv {N\choose 2}^{-1} \sum_{p<q} w_{pq},
+  \bar w \equiv {N\choose 2}^{-1} \sum_{p<q} w_{pq},
   &&
@@
   H_2^{(2)} = - \P_0 \V_2\p{\m w} \E \V_2\p{\m w} \P_0
-  = -\sum_{\Delta>0} \f1\Delta \P_0 \V_2\p{\m w_\Delta}^2 \P_0.
+  = -\sum_{\Delta>0} \f1\Delta \P_0 \V_2\p{\m w_\Delta}^2 \P_0,
 \end{align}
-The product $\P_0 \V_2\p{\m w_\Delta}^2 \P_0$ can be written in terms
-collective operators using techniques similar to those in Appendix
-\ref{sec:sym_prod}, which should yield a effective Hamiltonian of the
-form
-\begin{align}
-  H_2^{(2)} \simeq f\p{\m w_\Delta} \O\p{\v X,\v Y},
-\end{align}
-where $f\p{\m w_\Delta}$ is a scalar and $\O\p{\v X,\v Y}$ is a
-collective operator that is independent of $\m w_\Delta$.
+where the product $\P_0 \V_2\p{\m w_\Delta}^2 \P_0$ can be written in
+terms collective operators using techniques similar to those in
+Appendix \ref{sec:sym_prod}.
+
+% The product $\P_0 \V_2\p{\m w_\Delta}^2 \P_0$ can be written in
+% terms collective operators using techniques similar to those in
+% Appendix \ref{sec:sym_prod}, which should yield a effective
+% Hamiltonian of the form
+% \begin{align}
+%   H_2^{(2)} \simeq f\p{\m w_\Delta} \O\p{\v X,\v Y},
+% \end{align}
+% where $f\p{\m w_\Delta}$ is a scalar and $\O\p{\v X,\v Y}$ is a
+% collective operator that is independent of $\m w_\Delta$.
 
@@
   = \sum_{\substack{k\\p<q}} h_{pq} v_k \Pi_{pq} V_k \ket\psi,
-  \label{eq:H_V1_psi_start}
+  \label{eq:H_V1_psi_start_apndx}
 \end{align}
@@
 where we have used the facts that $h_{pq}=h_{qp}$ and $h_{pp}=0$.  The
-terms with $k\notin\set{p,q}$ in \eqref{eq:H_V1_psi_start} are then
+terms in \eqref{eq:H_V1_psi_start_apndx} with $k\notin\set{p,q}$ are
+then
 \begin{align}
@@
 where $\v h\oc\v v$ is an element-wise (Hadamard) product of $\v h$
-and $\v v$.  In the case of $k\in\set{p,q}$, we have
+and $\v v$.  The terms with $k\in\set{p,q}$, meanwhile, are
 \begin{align}
@@
   = \f12 \sum_{p\ne q} w_{pq} \P_0 X_p Y_q \P_0
-  = {N\choose 2} w \P_0 X_0 Y_1 \P_0,
+  = {N\choose 2} \bar w \P_0 X_0 Y_1 \P_0,
   \label{eq:PXYP_start}
@@
 \begin{align}
-  w \equiv {N\choose 2}^{-1} \sum_{p<q} w_{pq}
+  \bar w \equiv {N\choose 2}^{-1} \sum_{p<q} w_{pq}
 \end{align}
-is the mean value of the pair-wise coefficients in $\V_2$.  To
-simplify the operator content of \eqref{eq:PXYP_start}, we define
+is the mean value of the pair-wise couplings in $\V_2$.  To simplify
+the operator content of \eqref{eq:PXYP_start}, we define
 \begin{align}
@@
 \begin{align}
-  \P_0 \V_2 \P_0 = \f12 w \P_0 \p{\tilde Z -  Z } \P_0.
+  \P_0 \V_2 \P_0 = \f12 \bar w \P_0 \p{\tilde Z -  Z } \P_0.
 \end{align}

### 2019-11-05 collective_sun/multilevel_spin_notes.tex
added section on a periodic drive

@@
 \newcommand{\I}{\mathcal{I}}
+\newcommand{\J}{\mathcal{J}}
 \newcommand{\M}{\mathcal{M}}
@@
 combinations and products thereof.  Closure of $\M_\FS$ under
-collective the action of collective operators, and in particular
+collective the action of collective operators, and in particular under
 collective dynamics, is a straightforward consequence of the
-permutational symmetry obeyed by collective operators.
+permutational symmetry obeyed by these operators.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Periodic drive}
+
+We now consider the effect of applying periodic drive of the form
+\begin{align}
+  H_{\t{pd}}\p{t} = \sum_p \chi_p \omega \cos\p{\omega t} D_\z^{(p)},
+\end{align}
+where $\omega$ is a driving frequency, and $\chi_p$ is a (real)
+dimensionless driving amplitude for spin $p$, and
+$D_\z^{(p)} \equiv \xi_1 D_{1,0}^{(p)}$ with $\xi_1$ defined in
+\eqref{eq:scale_fac}.  The periodic drive $H_{\t{pd}}\p{t}$ can be
+implemented directly using external magnetic fields or driving lasers.
+Transition operators $T_{LM}$ transform under rotations generated by
+the drive operator $D_\z$ as
+\begin{align}
+  e^{i\phi D_\z} T_{LM} e^{-i\phi D_\z} = e^{iM\phi} T_{LM}.
+\end{align}
+If we move into the rotating frame of the drive, then the interaction
+Hamiltonian transforms as
+\begin{align}
+  H_{\t{int}} \to \tilde H_{\t{int}}
+  \equiv U_{\t{pd}}\p{t}^\dag H_{\t{int}} U_{\t{pd}}\p{t}
+\end{align}
+where
+\begin{align}
+  U_{\t{pd}}\p{t} \equiv \exp\sp{-i\int_0^td\tau\,H_{\t{pd}}\p{\tau}}
+  = \prod_p \exp\sp{-i \chi_p \sin\p{\omega t} D_\z^{(p)}}.
+\end{align}
+Expanding the interaction Hamiltonian $H_{\t{int}}$ in terms of
+transition operators $T_{LM}$, we find that
+\begin{align}
+  \tilde H_{\t{int}}
+  = -\f{u}{2} \sum_{p,q,L,M} e^{iM\p{\chi_q-\chi_p}\sin\p{\omega t}}
+  T_{LM}^{(p)\dag} T_{LM}^{(q)}
+  = -\f{u}{2} \sum_{p,q,L,M,n}
+  \J_n\p{M\sp{\chi_q-\chi_p}} e^{in\omega t}
+  T_{LM}^{(p)\dag} T_{LM}^{(q)},
+\end{align}
+where $\J_n$ is the $n$-th order Bessel function of the first kind.
+If the drive frequency $\omega$ is much larger than the interaction
+strength $u$, i.e.~$\omega\gg u$, then we can neglect oscillating
+terms with $n\ne0$, arriving at
+\begin{align}
+  \tilde H_{\t{int}} \approx -\f{u}{2} \sum_{p,q,L,M}
+  \J_0\p{M\sp{\chi_q-\chi_p}}
+  T_{LM}^{(p)\dag} T_{LM}^{(q)}.
+\end{align}
+
+\vspace{5cm}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Possible future directions}
@@
   \end{itemize}
-\item Classifying eigenstates of SU($n$)-symmetric interactions
 \item Implications of SOC

### 2019-11-08 collective_sun/sun_perturbation.tex
some changes to notation

@@
 \newcommand{\FS}{\text{FS}}
+\newcommand{\FSEQ}{=_{\text{FS}}}
 
@@
 To compute the effective Hamiltonian induced by the single-body
-perturbation $\V_1$, we expand
+perturbation $\V_1$, we use the permutational symmetry of the
+ground-state manifold to simplify
 \begin{align}
-  H_1^{(1)} = \v v \c \P_0 \v V \P_0,
-\end{align}
-and use the permutational symmetry of the ground-state manifold to
-simplify
-\begin{align}
-  H_1^{(1)} =  \sum_p v_p \P_0 V_0 \P_0 \simeq \bar v V,
-  &&
-  \bar v \equiv \EE_p\sp{v_p} = \f1N \sum_p v_p,
-  &&
-  V \equiv \sum_p V_p,
+  H_1^{(1)} = \sum_p v_p \P_0 V_p \P_0
+  = \sum_p v_p \P_0 V_0 \P_0 \FSEQ \bar v V,
   \label{eq:H_1_1}
 \end{align}
-where $\simeq$ denotes equality up to a restriction to the
-ground-state manifold $\M_0$, and $\EE_p\sp{x_p}$ denotes the mean
-value of $x_p$.  In order to compute the second order effective
-Hamiltonian $H_1^{(2)}$, we first choose an arbitrary fully symmetric
-state $\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_V1_psi}):
+where $\FSEQ$ denotes equality under a restriction to the fully
+symmetric manifold $\M_0$, and
+\begin{align}
+  \bar v \equiv \f1N \sum_p v_p,
+  &&
+  V \equiv \sum_p V_p.
+\end{align}
+In order to compute the second order effective Hamiltonian
+$H_1^{(2)}$, we first choose an arbitrary fully symmetric state
+$\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_V1_psi}):
 \begin{align}
@@
 definite interaction energy $\Delta$ above that of the fully symmetric
-manifold.  We can therefore simplify
+manifold $\M_0$.  We can therefore simplify
 \begin{align}
   H_1^{(2)}
-  = -\P_0 \V_1\p{\v v} \E \V_1\p{\v v} \P_0
+  = - \P_0 \V_1\p{\v v} \E \V_1\p{\v v} \P_0
   = - \sum_{\Delta>0} \f1{\Delta}
@@
 where the product $\P_0 \V_1\p{\v v_\Delta}^2 \P_0$ is worked out in
-Appendix \ref{sec:PVVP}, giving us
+Appendix \ref{sec:PVVP}.  In total, we find
 \begin{align}
-  H_1^{(2)} \simeq \f{\tilde v}{N-1} \p{V^2 - N W},
+  H_1^{(2)} \FSEQ \f{\tilde v}{N-1} \p{V^2 - N W},
   &&
-  \tilde v \equiv \sum_{\Delta>0} \f{\var\p{\v v_\Delta}}{\Delta}.
+  \tilde v \equiv \sum_{\Delta>0} \f{\var\p{\v v_\Delta}}{\Delta},
   \label{eq:H_1_2}
 \end{align}
-Here $\simeq$ denotes equality up to a restriction to the fully
+where $\FSEQ$ denotes equality under a restriction to the fully
 symmetric manifold $\M_0$, and
@@
 \end{align}
-with $\EE_p\sp{x_p}\equiv\sum_px_p/N$ the mean of $\v x$.
+with $\EE_p\sp{x_p}\equiv\sum_px_p/N$ the mean value in $\v x$.
 
@@
 symmetric manifold $\M_0$ by the two-body perturbation $\V_2$ in
-\eqref{eq:perturbations} can be computed using the permutational
-symmetry of $\M_0$ (see Appendix \ref{sec:PXYP}):
+\eqref{eq:perturbations} is simply the restriction of $\V_2$ onto
+$\M_0$ (see Appendix \ref{sec:PXYP}):
 \begin{align}
   H_2^{(1)}
-  = \f12 \P_0 \v X\c\m w\c\v Y\P_0
-  \simeq \f12 \bar w \p{\tilde Z - Z},
+  = \P_0 \V_2 \P_0
+  \FSEQ \f12 \bar w \p{\tilde Z - Z},
 \end{align}
@@
 
-% The product $\P_0 \V_2\p{\m w_\Delta}^2 \P_0$ can be written in
-% terms collective operators using techniques similar to those in
-% Appendix \ref{sec:sym_prod}, which should yield a effective
-% Hamiltonian of the form
-% \begin{align}
-%   H_2^{(2)} \simeq f\p{\m w_\Delta} \O\p{\v X,\v Y},
-% \end{align}
-% where $f\p{\m w_\Delta}$ is a scalar and $\O\p{\v X,\v Y}$ is a
-% collective operator that is independent of $\m w_\Delta$.
-
 \appendix
@@
 on the fully symmetric manifold $\M_0$ by the perturbation $\V_1$ in
-\eqref{eq:perturbations} requires us to simplify the product
+\eqref{eq:perturbations} requires us to simplify
 \begin{align}
   \P_0 \V_1\p{\v v}^2 \P_0
-  = \sum_{p,q} v_p v_q \P_0 V_p V_q \P_0
-  = \sum_p v_p^2 \P_0 V_p^2 \P_0
-  + \sum_{p\ne q} v_p v_q \P_0 V_p V_q \P_0,
+  \FSEQ \sum_{p,q} v_p v_q V_p V_q
+  = \sum_p v_p^2 V_p^2
+  + \sum_{p\ne q} v_p v_q V_p V_q,
   \label{eq:PVVP_start}
 \end{align}
-where the coefficients $\v v$ satisfy \eqref{eq:cond_1} with
-eigenvalue $\Delta\ne0$, which implies that $\v v$ is mean-zero.  The
-first sum on the right of \eqref{eq:PVVP_start} is then
+where $\FSEQ$ denotes equality under a restriction to the fully
+symmetric manifold $\M_0$; and the coefficients $\v v$ satisfy
+\eqref{eq:cond_1} with eigenvalue $\Delta\ne0$, which implies that
+$\v v$ is mean-zero.  The first sum on the right of
+\eqref{eq:PVVP_start} is then
 \begin{align}
-  \sum_p v_p^2 \P_0 V_p^2 \P_0
-  = \sum_p v_p^2 \P_0 V_0^2 \P_0
-  = \var\p{\v v} \P_0 W \P_0,
+  \sum_p v_p^2 V_p^2
+  \FSEQ \sum_p v_p^2 V_0^2
+  \FSEQ \var\p{\v v} W,
   \label{eq:PVVP_eq}
@@
 \end{align}
-The second sum in \eqref{eq:PVVP_start} is
+The second sum on the right of \eqref{eq:PVVP_start} is
 \begin{align}
-  \sum_{p\ne q} v_p v_q \P_0 V_p V_q \P_0
-  = \sum_{p\ne q} v_p v_q \P_0 V_0 V_1 \P_0,
+  \sum_{p\ne q} v_p v_q V_p V_q
+  \FSEQ \sum_{p\ne q} v_p v_q V_0 V_1,
   \label{eq:PVVP_neq_start}
@@
 \begin{align}
-  \P_0 V^2 \P_0
-  = \sum_p \P_0 V_p^2 \P_0
-  + \sum_{p\ne q} \P_0 V_p V_q \P_0
-  = \P_0 W \P_0 + N \p{N-1} \P_0 V_0 V_1 \P_0,
+  V^2
+  = \sum_p V_p^2
+  + \sum_{p\ne q} V_p V_q
+  \FSEQ W + N \p{N-1} V_0 V_1,
   \label{eq:PVVP_neq_ops}
@@
 \begin{align}
-  \sum_{p\ne q} v_p v_q \P_0 V_p V_q \P_0
-  = - \var\p{\v v} N \P_0 V_0 V_1 \P_0
-  = -\f{\var\p{\v v}}{N-1} \P_0 \p{V^2 - W} \P_0
+  \sum_{p\ne q} v_p v_q V_p V_q
+  \FSEQ - \var\p{\v v} N V_0 V_1
+  \FSEQ -\f{\var\p{\v v}}{N-1} \p{V^2 - W}
   \label{eq:PVVP_neq}
@@
 \begin{align}
-  \P_0 \V_1\p{\v v}^2 \P_0
-  = -\f{\var\p{\v v}}{N-1} \P_0 \p{V^2 - N W} \P_0,
+  \V_1\p{\v v}^2
+  \FSEQ -\f{\var\p{\v v}}{N-1} \p{V^2 - N W}.
 \end{align}
@@
 on the fully symmetric manifold $\M_0$ by the perturbation $\V_2$ in
-\eqref{eq:perturbations} requires us to simplify the product
+\eqref{eq:perturbations} requires us to simplify
 \begin{align}
   \P_0 \V_2 \P_0
-  = \f12 \sum_{p\ne q} w_{pq} \P_0 X_p Y_q \P_0
-  = {N\choose 2} \bar w \P_0 X_0 Y_1 \P_0,
+  \FSEQ \f12 \sum_{p\ne q} w_{pq} X_p Y_q
+  = {N\choose 2} \bar w X_0 Y_1,
   \label{eq:PXYP_start}
 \end{align}
-where ${N\choose2}=N\p{N-1}/2$ and
+where $\FSEQ$ denotes equality under a restriction to the fully
+symmetric manifold $\M_0$, ${N\choose2}=N\p{N-1}/2$ is a binomial
+coefficient, and
 \begin{align}
@@
 \begin{align}
-  \P_0 \tilde Z \P_0 = \P_0 Z \P_0 + N\p{N-1} \P_0 X_0 Y_1 \P_0.
+  \tilde Z \FSEQ Z + N\p{N-1} X_0 Y_1.
 \end{align}
-Substituting the product $\P_0 X_0 Y_1 \P_0$ back into
-\eqref{eq:PXYP_start}, we thus find that
+Substituting the product $X_0 Y_1$ back into \eqref{eq:PXYP_start}
+gives us
 \begin{align}
-  \P_0 \V_2 \P_0 = \f12 \bar w \P_0 \p{\tilde Z -  Z } \P_0.
+  \V_2 \FSEQ \f12 \bar w \p{\tilde Z -  Z }.
 \end{align}

### 2019-11-09 collective_sun/sun_perturbation.tex
overhauled some notation that will allow finishing the calculation of the second-order effect of two-body perturabtions

@@
 \newcommand{\X}{\mathcal{X}}
+\newcommand{\Y}{\mathcal{Y}}
 \newcommand{\Z}{\mathcal{Z}}
@@
 \begin{align}
-  \V_1 \equiv \sum_p v_p V_p = \v v\c\v V,
+  \V_1 \equiv \sum_p v_p X_p = \v v\c\v X,
   &&
@@
 \end{align}
-where $V_p,X_p,Y_p$ are trace-zero operators on spin $p$, with
-e.g.~$V_p$ related to $V_q$ by a permutation of spins $p$ and $q$;
-$\v V,\v X,\v Y$ are respectively vectors of all $V_p,X_p,Y_p$; $\v v$
-and $\m w$ are respectively a vector and matrix of all $v_p,w_{pq}$;
-we enforce $w_{pq}=w_{qp}$ and $w_{pp}=0$; and we use a dot ($\c$) to
-denote an inner product between vectors and matrices as appropriate.
+where $X_p,Y_p$ are trace-zero operators on spin $p$, with e.g.~$X_p$
+related to $X_q$ by a permutation of spins $p$ and $q$; $\v X,\v Y$
+are respectively vectors of all $X_p,Y_p$; $\v v$ and $\m w$ are
+respectively a vector and matrix of all $v_p,w_{pq}$; we enforce
+$w_{pq}=w_{qp}$ and $w_{pp}=0$; and we use a dot ($\c$) to denote an
+inner product between vectors and matrices as appropriate.
 
@@
 \begin{align}
-  H_1^{(1)} = \sum_p v_p \P_0 V_p \P_0
-  = \sum_p v_p \P_0 V_0 \P_0 \FSEQ \bar v V,
+  H_1^{(1)} = \sum_p v_p \P_0 X_p \P_0
+  = \sum_p v_p \P_0 X_0 \P_0 \FSEQ \bar v \X,
   \label{eq:H_1_1}
@@
   &&
-  V \equiv \sum_p V_p.
+  \X \equiv \sum_p X_p.
 \end{align}
@@
   H_0 \V_1 \ket\psi
-  = E_0 \V_1 \ket\psi
-  + \p{\v v\c\m h - \v v\oc\v h}\c \v V \ket\psi,
+  &= E_0 \V_1 \ket\psi
+  + \p{\sum_{p,q} h_{pq} v_p X_q - \sum_q v_q h_q X_q} \ket\psi \\
+  &= E_0 \V_1 \ket\psi
+  + \v v\c \p{\m h - \diag\v h}\c \v X \ket\psi,
   \label{eq:H_V1_psi_start}
@@
 interaction Hamiltonian $H_0$ in \eqref{eq:ints}; $\v h$ is a vector
-of the net interaction energies $h_p\equiv\sum_qh_{pq}$ associated
-with individual spins $p$; and $\v v\oc\v h$ is the element-wise
-(Hadamard) product of $\v v$ and $\v h$, i.e.~a vector with components
-$v_kh_k$.  We thus find that the vector $\V_1\ket\psi$ is an
-eigenvector of the interaction Hamiltonian $H_0$ if the coefficients
-$\v v$ satisfy
+of the net interaction energies $h_q\equiv\sum_ph_{pq}$ associated
+with individual spins $p$; and $\diag\v h$ is a matrix with $\v h$ on
+the diagonal and zeros everywhere else.  We thus find that the vector
+$\V_1\ket\psi$ is an eigenvector of the interaction Hamiltonian $H_0$
+if the coefficients $\v v$ satisfy
 \begin{align}
-  \m h \c \v v - \v h\oc\v v
-  = \p{\m h - \diag\v h}\c\v v
-  = \Delta\p{\v v} \v v,
+  \p{\m h - \diag\v h}\c\v v = \Delta\p{\v v} \v v.
   \label{eq:cond_1}
-\end{align}
-where $\diag\v h$ is a matrix with $\v h$ on the diagonal and zeroes
-everywhere else; and $\Delta\p{\v v}$ is some constant that may depend
-on $\v v$.  If the interaction Hamiltonian $H_0$ is translationally
-invariant, then all $h_p=h\equiv\EE_q\sp{h_q}$ and
-$\v h\oc\v v=h\v v$, so the condition in \eqref{eq:cond_1} simplifies
-to
-\begin{align}
-  \m h \c\v v = \sp{\Delta\p{\v v}+h} \v v.
 \end{align}
@@
 $\v v_\Delta$, and is entirely independent of the state $\ket\psi$ or
-choice of single-spin trace-zero operator $V_p$ used to build $\V_1$
+choice of single-spin trace-zero operator $X_p$ used to build $\V_1$
 in \eqref{eq:perturbations}.  Finding operators $\V_1\p{\v v_\Delta}$
@@
 where the product $\P_0 \V_1\p{\v v_\Delta}^2 \P_0$ is worked out in
-Appendix \ref{sec:PVVP}.  In total, we find
+Appendix \ref{sec:PXXP}.  In total, we find
 \begin{align}
-  H_1^{(2)} \FSEQ \f{\tilde v}{N-1} \p{V^2 - N W},
+  H_1^{(2)} \FSEQ \f{\tilde v}{N-1} \p{\X^2 - N \X^{(2)}},
   &&
@@
 \begin{align}
+  \X^{(2)} \equiv \sum_p X_p^2,
+  &&
   \var\p{\v x} \equiv \EE_p\sp{\p{x_p-\EE_q\sp{x_q}}^2}
   = \EE_p\sp{x_p^2} - \EE_q\sp{x_q}^2,
-  &&
-  W \equiv \sum_p V_p^2,
 \end{align}
@@
   = \P_0 \V_2 \P_0
-  \FSEQ \f12 \bar w \p{\tilde Z - Z},
+  \FSEQ \f12 \bar w \p{\X\odot\Y - \O_{X,Y}},
 \end{align}
-where in terms of the anti-commutator $\sp{A,B}_+\equiv AB+BA$ and
-collective operators $X\equiv\sum_p X_p$ and $Y\equiv\sum_p Y_p$, we
-define
+where $A\odot B\equiv\p{AB+BA}/2$ is a symmetrized product of $A$ and
+$B$; and
 \begin{align}
@@
   &&
-  \tilde Z \equiv \f12 \sp{X,Y}_+,
-  &&
-  Z \equiv \sum_p \f12 \sp{X_p, Y_p}_+.
+  \O_{X,Y} \equiv \sum_p X_p \odot Y_p.
 \end{align}
@@
   H_0 \V_2 \ket\psi
-  = E_0 \V_2 \ket\psi
+  &= E_0 \V_2 \ket\psi
+  + \sum_{k,\ell} \bar h_{k\ell} w_{k\ell} X_k Y_\ell \ket\psi
+  + \sum_{k,\ell,q} h_{kq} w_{q\ell} X_k Y_\ell \ket\psi
+  - \sum_{k,\ell} w_{k\ell} h_{k\ell} X_k \odot Y_k \ket\psi \\
+  &= E_0 \V_2 \ket\psi
   + \v X \c \p{\bar{\m h}\oc\m w + \m h\c\m w} \c \v Y \ket\psi
-  - \v\1 \c \p{\m h\oc\m w} \c \v Z \ket\psi,
+  - \v\1 \c \p{\m h\oc\m w} \c \v O_{X,Y} \ket\psi,
   \label{eq:H_V2_psi}
 \end{align}
-where $\v\1$ is an $N$-component vector of ones;
+where $\m w,\m h,\bar{\m h}$ are matrices of all
+$w_{k\ell},h_{k\ell},\bar h_{k\ell}$, with
 \begin{align}
-  \bar{\m h}
-  \equiv \m h - \f12\p{\v h\otimes\v\1 + \v\1\otimes\v h}
+  \bar h_{k\ell} \equiv h_{k\ell} - \f12\p{h_k + h_\ell},
+  &&
+  h_k \equiv \sum_\ell h_{k\ell};
 \end{align}
-is a modified interaction matrix with entries $h_{pq}-\p{h_p+h_q}/2$;
-and $\m x\oc\m y$ is an element-wise (Hadamard) product of $\m x$ and
-$\m y$, with entries $x_{pq}y_{pq}$.  We thus find that the vector
-$\V_2\ket\psi$ is an eigenvector of the interaction Hamiltonian $H_0$
-if the matrix $\m w$ of all $w_{pq}$ satisfies
+$\m x\oc\m y$ is an element-wise (Hadamard) product of $\m x$ and
+$\m y$, i.e.~with matrix elements $x_{k\ell} y_{k\ell}$;
+$\v\1\equiv\p{1,1,1,\cdots}$ is an $N$-component vector of ones; and
+$\v O_{X,Y}$ is a vector with components $\p{X_kY_k+Y_kX_k}/2$, i.e.
+\begin{align}
+  \v O_{X,Y} \equiv \f12\p{\v X\oc\v Y + \v Y\oc\v X}.
+\end{align}
+We thus find that the vector $\V_2\ket\psi$ is an eigenvector of the
+interaction Hamiltonian $H_0$ if the coefficient matrix $\m w$ of all
+$w_{pq}$ satisfies
 \begin{align}
@@
 components of the condition in \eqref{eq:cond_2} automatically enforce
-that the last term ($\sim\v Z$) in \eqref{eq:H_V2_psi} vanishes.
+that the last term ($\sim\v O_{X,Y}$) in \eqref{eq:H_V2_psi} vanishes.
 
@@
 \begin{align}
-  \check{\m h}
+  \tilde{\m h}
   \equiv \sum_{p,q} \bar h_{pq} \op{pq}
@@
 \begin{align}
-  \check{\m h} \c \v{\m w} = \f12 \Delta\p{\m w} \v{\m w}.
+  \tilde{\m h} \c \v{\m w} = \f12 \Delta\p{\m w} \v{\m w}.
   \label{eq:cond_2_eig}
 \end{align}
-The matrix $\check{\m h}$ in \eqref{eq:h_super_mat} can be written in
+The matrix $\tilde{\m h}$ in \eqref{eq:h_super_mat} can be written in
 the block-diagonal form
 \begin{align}
-  \check{\m h} = \sum_\ell \m h_\ell \otimes \op{\ell},
+  \tilde{\m h} = \sum_\ell \m h_\ell \otimes \op{\ell},
   &&
@@
 eigenvalues $\Delta\p{\m w}/2$.  ``Total'' eigenvectors $\v{\m w}$ of
-$\check{\m h}$ are then built from eigenvectors $\v w_\ell$ within
+$\tilde{\m h}$ are then built from eigenvectors $\v w_\ell$ within
 each block $\m h_\ell$ that share the same eigenvalue
@@
   H_0 \V_1 \ket\psi
-  = \sum_{\substack{k\\p<q}} h_{pq} v_k \Pi_{pq} V_k \ket\psi,
+  = \sum_{\substack{k\\p<q}} h_{pq} v_k \Pi_{pq} X_k \ket\psi,
   \label{eq:H_V1_psi_start_apndx}
@@
 $k\notin\set{p,q}$.  In the case of $k\notin\set{p,q}$, the
-permutation operator $\Pi_{pq}$ commutes with $V_k$ and annihilates on
+permutation operator $\Pi_{pq}$ commutes with $X_k$ and annihilates on
 $\ket\psi$, leaving us the sum
@@
   \sum_{\substack{p<q\\k\notin\set{p,q}}}
-  h_{pq} v_k \Pi_{pq} V_k \ket\psi
-  = \sum_k \p{E_0 - h_k} v_k V_k \ket\psi
-  = E_0 \V_1 \ket\psi - \v h\oc\v v \c\v V \ket\psi,
+  h_{pq} v_k \Pi_{pq} X_k \ket\psi
+  = \sum_k \p{E_0 - h_k} v_k X_k \ket\psi
+  = E_0 \V_1 \ket\psi - \sum_k h_k v_k X_k \ket\psi
 \end{align}
-where $\v h\oc\v v$ is an element-wise (Hadamard) product of $\v h$
-and $\v v$.  The terms with $k\in\set{p,q}$, meanwhile, are
+while the terms with $k\in\set{p,q}$ are
 \begin{align}
   \sum_{\substack{p<q\\k\in\set{p,q}}}
-  h_{pq} v_k \Pi_{pq} V_k \ket\psi
-  = \sum_{k<q} h_{kq} v_k V_q \ket\psi
-  + \sum_{p<k} h_{pk} v_k V_p \ket\psi
-  = \sum_{k,q} h_{kq} v_k V_q \ket\psi
-  = \v v\c\m h\c\v V \ket\psi,
+  h_{pq} v_k \Pi_{pq} X_k \ket\psi
+  = \sum_{k<q} h_{kq} v_k X_q \ket\psi
+  + \sum_{p<k} h_{pk} v_k X_p \ket\psi
+  = \sum_{k,q} h_{kq} v_k X_q \ket\psi
 \end{align}
@@
   = E_0 \V_1 \ket\psi
-  + \p{\v v\c\m h - \v v\oc\v h}\c \v V \ket\psi.
+  + \p{\sum_{p,q} h_{pq} v_p X_q - \sum_q h_q v_q X_q} \ket\psi.
 \end{align}
@@
   - \sum_{\substack{p\in\set{k,\ell}\\q\notin\set{k,\ell}}} h_{pq}
-  = E_0 + 2 h_{k\ell} - h_k - h_\ell.
+  = E_0 + 2 h_{k\ell} - h_k - h_\ell
+  = E_0 + 2 \bar h_{k\ell},
+\end{align}
+where
+\begin{align}
+  \bar h_{k\ell} \equiv h_{k\ell} - \f12\p{h_k + h_\ell}.
 \end{align}
@@
 \begin{align}
-  E_0 \V_2 \ket\psi + \sum_{k,\ell}
-  \sp{h_{k\ell} - \f12\p{h_k+h_\ell}} w_{k\ell} X_k Y_\ell \ket\psi
-  = E_0 \V_2 \ket\psi + \v X \c \p{\bar{\m h}\oc\m w} \c \v Y \ket\psi,
+  E_0 \V_2 \ket\psi
+  + \sum_{k,\ell} \bar h_{k\ell} w_{k\ell} X_k Y_\ell \ket\psi.
 \end{align}
-where $\bar{\m h}\oc\m w$ is an element-wise (Hadamard) product
-between $\bar{\m h}$ and $\m w$,
-\begin{align}
-  \bar{\m h} \equiv \m h - \f12\p{\v h\otimes\v\1 + \v\1\otimes\v h},
-\end{align}
-and $\v\1$ is a vector of ones.  The terms in
-\eqref{eq:H_V2_psi_start} with only one of $p,q\in\set{k,\ell}$
-simplify to
+The terms in \eqref{eq:H_V2_psi_start} with only one of
+$p,q\in\set{k,\ell}$ simplify to
 \begin{align}
@@
   \p{h_{kq} X_q Y_\ell + h_{\ell q} X_k Y_q} \ket\psi
-  &= \sum_{k,\ell,q} w_{k\ell} h_{kq} X_q Y_\ell \ket\psi
-  - \sum_{k,\ell} w_{k\ell} h_{k\ell} Z_k \ket\psi \\
-  &= \v X \c\m h \c\m w\c\v Y \ket\psi
-  - \v\1 \c \p{\m h\oc\m w}\c \v Z \ket\psi,
+  = \sum_{k,\ell,q} w_{k\ell} h_{kq} X_q Y_\ell \ket\psi
+  - \sum_{k,\ell} w_{k\ell} h_{k\ell} X_k \odot Y_k \ket\psi
 \end{align}
-where
-\begin{align}
-  Z_k \equiv \f12 \sp{X_k, Y_k}_+,
-\end{align}
-and $\v Z$ is a vector of all $Z_p$.  In total,
+where $A\odot B \equiv \p{AB+BA}/2$ is a symmetrized product.  In
+total,
 \begin{align}
@@
   = E_0 \V_2 \ket\psi
-  + \v X \c \p{\bar{\m h}\oc\m w + \m h\c\m w} \c \v Y \ket\psi
-  - \v\1 \c \p{\m h\oc\m w}\c \v Z \ket\psi.
+  + \sum_{k,\ell} \bar h_{k\ell} w_{k\ell} X_k Y_\ell \ket\psi
+  + \sum_{k,\ell,q} h_{kq} w_{q\ell} X_k Y_\ell \ket\psi
+  - \sum_{k,\ell} w_{k\ell} h_{k\ell} X_k \odot Y_k \ket\psi.
 \end{align}
@@
 \subsection{Single-body product}
-\label{sec:PVVP}
+\label{sec:PXXP}
 
@@
   \P_0 \V_1\p{\v v}^2 \P_0
-  \FSEQ \sum_{p,q} v_p v_q V_p V_q
-  = \sum_p v_p^2 V_p^2
-  + \sum_{p\ne q} v_p v_q V_p V_q,
-  \label{eq:PVVP_start}
+  \FSEQ \sum_{p,q} v_p v_q X_p X_q
+  = \sum_p v_p^2 X_p^2
+  + \sum_{p\ne q} v_p v_q X_p X_q,
+  \label{eq:PXXP_start}
 \end{align}
@@
 $\v v$ is mean-zero.  The first sum on the right of
-\eqref{eq:PVVP_start} is then
+\eqref{eq:PXXP_start} is then
 \begin{align}
-  \sum_p v_p^2 V_p^2
-  \FSEQ \sum_p v_p^2 V_0^2
-  \FSEQ \var\p{\v v} W,
-  \label{eq:PVVP_eq}
+  \sum_p v_p^2 X_p^2
+  \FSEQ \sum_p v_p^2 X_0^2
+  \FSEQ \var\p{\v v} \X^{(2)},
+  \label{eq:PXXP_eq}
 \end{align}
@@
   &&
-  W \equiv \sum_p V_p^2.
+  \X^{(2)} \equiv \sum_p X_p^2.
 \end{align}
-The second sum on the right of \eqref{eq:PVVP_start} is
+The second sum on the right of \eqref{eq:PXXP_start} is
 \begin{align}
-  \sum_{p\ne q} v_p v_q V_p V_q
-  \FSEQ \sum_{p\ne q} v_p v_q V_0 V_1,
-  \label{eq:PVVP_neq_start}
+  \sum_{p\ne q} v_p v_q X_p X_q
+  \FSEQ \sum_{p\ne q} v_p v_q X_0 X_1,
+  \label{eq:PXXP_neq_start}
 \end{align}
-where the sum over coefficients in \eqref{eq:PVVP_neq_start} can be
+where the sum over coefficients in \eqref{eq:PXXP_neq_start} can be
 simplified to
@@
 In order to simplify the product of operators in
-\eqref{eq:PVVP_neq_start}, we expand
+\eqref{eq:PXXP_neq_start}, we expand
 \begin{align}
-  V^2
-  = \sum_p V_p^2
-  + \sum_{p\ne q} V_p V_q
-  \FSEQ W + N \p{N-1} V_0 V_1,
-  \label{eq:PVVP_neq_ops}
+  \X^2
+  = \sum_p X_p^2
+  + \sum_{p\ne q} X_p X_q
+  \FSEQ \X^{(2)} + N \p{N-1} X_0 X_1,
+  \label{eq:PXXP_neq_ops}
 \end{align}
@@
 \begin{align}
-  \sum_{p\ne q} v_p v_q V_p V_q
-  \FSEQ - \var\p{\v v} N V_0 V_1
-  \FSEQ -\f{\var\p{\v v}}{N-1} \p{V^2 - W}
-  \label{eq:PVVP_neq}
+  \sum_{p\ne q} v_p v_q X_p X_q
+  \FSEQ - N \var\p{\v v} X_0 X_1
+  \FSEQ -\f{\var\p{\v v}}{N-1} \p{\X^2 - \X^{(2)}}
+  \label{eq:PXXP_neq}
 \end{align}
@@
   \V_1\p{\v v}^2
-  \FSEQ -\f{\var\p{\v v}}{N-1} \p{V^2 - N W}.
+  \FSEQ -\f{\var\p{\v v}}{N-1} \p{\X^2 - N \X^{(2)}}.
 \end{align}
@@
 \begin{align}
-  X \equiv \sum_p X_p,
+  \X \equiv \sum_p X_p,
   &&
-  Y \equiv \sum_p Y_p,
+  \Y \equiv \sum_p Y_p,
   &&
-  Z_p \equiv \f12\sp{X_p, Y_p}_+,
+  A \odot B \equiv \f12\p{AB+BA},
   &&
-  Z \equiv \sum_p Z_p,
-  &&
-  \tilde Z \equiv \f12 \sp{X, Y}_+,
+  \O_{X,Y} \equiv \sum_p X_p \odot Y_p,
 \end{align}
@@
 \begin{align}
-  \tilde Z \FSEQ Z + N\p{N-1} X_0 Y_1.
+  \X\odot\Y \FSEQ \O_{X,Y} + N\p{N-1} X_0 Y_1.
 \end{align}
@@
 \begin{align}
-  \V_2 \FSEQ \f12 \bar w \p{\tilde Z -  Z }.
+  \V_2 \FSEQ \f12 \bar w \p{\X\odot\Y -  \O_{X,Y}}.
 \end{align}

### 2019-11-11 collective_sun/sun_perturbation.tex
added part of calculation for second-order effect of two-body perturbation

@@
 \newcommand{\FS}{\text{FS}}
-\newcommand{\FSEQ}{=_{\text{FS}}}
+\newcommand{\EQFS}{=_{\text{FS}}}
 
@@
   H_1^{(1)} = \sum_p v_p \P_0 X_p \P_0
-  = \sum_p v_p \P_0 X_0 \P_0 \FSEQ \bar v \X,
+  = \sum_p v_p \P_0 X_0 \P_0 \EQFS \bar v \X,
   \label{eq:H_1_1}
 \end{align}
-where $\FSEQ$ denotes equality under a restriction to the fully
+where $\EQFS$ denotes equality under a restriction to the fully
 symmetric manifold $\M_0$, and
@@
   + \v v\c \p{\m h - \diag\v h}\c \v X \ket\psi,
-  \label{eq:H_V1_psi_start}
+  \label{eq:H_V1_psi}
 \end{align}
@@
   = \p{E_0 + \Delta} \V_1\p{\v v_\Delta} \ket\psi.
-  \label{eq:H_V1_psi}
 \end{align}
@@
 \begin{align}
-  H_1^{(2)} \FSEQ \f{\tilde v}{N-1} \p{\X^2 - N \X^{(2)}},
-  &&
-  \tilde v \equiv \sum_{\Delta>0} \f{\var\p{\v v_\Delta}}{\Delta},
+  H_1^{(2)} \EQFS \tilde v \p{\X^2 - N \X^{(2)}},
   \label{eq:H_1_2}
 \end{align}
-where $\FSEQ$ denotes equality under a restriction to the fully
+where $\EQFS$ denotes equality under a restriction to the fully
 symmetric manifold $\M_0$, and
 \begin{align}
-  \X^{(2)} \equiv \sum_p X_p^2,
+  \tilde v \equiv \f1{N\p{N-1}}
+  \sum_{\Delta>0} \f{\v v_\Delta\c\v v_\Delta}{\Delta},
   &&
-  \var\p{\v x} \equiv \EE_p\sp{\p{x_p-\EE_q\sp{x_q}}^2}
-  = \EE_p\sp{x_p^2} - \EE_q\sp{x_q}^2,
+  \X^{(2)} \equiv \sum_p X_p^2.
 \end{align}
-with $\EE_p\sp{x_p}\equiv\sum_px_p/N$ the mean value in $\v x$.
+As the vectors $\v v_\Delta$ with $\Delta>0$ are mean-zero, we can
+equivalently write
+\begin{align}
+  \tilde v = \f1{N-1}
+  \sum_{\Delta>0} \f{\var\p{\v v_\Delta}}{\Delta},
+\end{align}
+where $\var\p{\v x}$ is the variance of the entries in $\v x$.
 
@@
   = \P_0 \V_2 \P_0
-  \FSEQ \f12 \bar w \p{\X\odot\Y - \O_{X,Y}},
+  \EQFS \f12 \bar w \p{\X\odot\Y - \Z},
 \end{align}
@@
   &&
-  \O_{X,Y} \equiv \sum_p X_p \odot Y_p.
+  \Z \equiv \sum_p Z_p,
+  &&
+  Z_p \equiv X_p \odot Y_p.
 \end{align}
@@
   + \sum_{k,\ell,q} h_{kq} w_{q\ell} X_k Y_\ell \ket\psi
-  - \sum_{k,\ell} w_{k\ell} h_{k\ell} X_k \odot Y_k \ket\psi \\
+  - \sum_{k,\ell} w_{k\ell} h_{k\ell} Z_k \ket\psi \\
   &= E_0 \V_2 \ket\psi
   + \v X \c \p{\bar{\m h}\oc\m w + \m h\c\m w} \c \v Y \ket\psi
-  - \v\1 \c \p{\m h\oc\m w} \c \v O_{X,Y} \ket\psi,
+  - \v\1 \c \p{\m h\oc\m w} \c \v Z \ket\psi,
   \label{eq:H_V2_psi}
@@
 $\v\1\equiv\p{1,1,1,\cdots}$ is an $N$-component vector of ones; and
-$\v O_{X,Y}$ is a vector with components $\p{X_kY_k+Y_kX_k}/2$, i.e.
-\begin{align}
-  \v O_{X,Y} \equiv \f12\p{\v X\oc\v Y + \v Y\oc\v X}.
-\end{align}
-We thus find that the vector $\V_2\ket\psi$ is an eigenvector of the
-interaction Hamiltonian $H_0$ if the coefficient matrix $\m w$ of all
-$w_{pq}$ satisfies
+$\v Z$ is a vector of all $Z_k$.  We thus find that the vector
+$\V_2\ket\psi$ is an eigenvector of the interaction Hamiltonian $H_0$
+if the coefficient matrix $\m w$ of all $w_{pq}$ satisfies
 \begin{align}
@@
 components of the condition in \eqref{eq:cond_2} automatically enforce
-that the last term ($\sim\v O_{X,Y}$) in \eqref{eq:H_V2_psi} vanishes.
+that the last term ($\sim\v Z$) in \eqref{eq:H_V2_psi} vanishes.
 
@@
 \begin{align}
-  \m h_\ell \c \v{\m w}_\ell = \f12 \Delta\p{\m w} \v w_\ell,
+  \m h_\ell \c \v w_\ell = \f12 \Delta\p{\m w} \v w_\ell,
   &&
-  \v w_\ell \equiv \sum_p w_{p\ell} \ket{p}.
+  \v w_\ell \equiv \sum_p w_{p\ell} \ket{p},
   \label{eq:cond_2_eig_block}
 \end{align}
-We can (numerically) diagonalize each of the blocks $\m h_\ell$, and
-classify their eigenvectors $\v w_\ell$ by the corresponding
-eigenvalues $\Delta\p{\m w}/2$.  ``Total'' eigenvectors $\v{\m w}$ of
-$\tilde{\m h}$ are then built from eigenvectors $\v w_\ell$ within
-each block $\m h_\ell$ that share the same eigenvalue
-$\Delta\p{\m w}/2$.  The task of constructing eigenvectors
+where $\v w_\ell$ is essentially the $\ell$-th row or column of
+$\m w$.  We can (numerically) diagonalize each of the blocks
+$\m h_\ell$, and classify their eigenvectors $\v w_\ell$ by the
+corresponding eigenvalues $\Delta\p{\m w}/2$.  ``Total'' eigenvectors
+$\v{\m w}$ of $\tilde{\m h}$ are then built from eigenvectors
+$\v w_\ell$ within each block $\m h_\ell$ that share the same
+eigenvalue $\Delta\p{\m w}/2$.  The task of constructing eigenvectors
 $\V_2\p{\m w_\Delta}\ket\psi$ of $H_0$ and determining their energy
@@
   = \p{E_0 + \Delta} \V_2\p{\m w_\Delta} \ket\psi,
-  \label{eq:H_V1_psi}
 \end{align}
@@
 terms collective operators using techniques similar to those in
-Appendix \ref{sec:sym_prod}.
+Appendix \ref{sec:sym_prod}.  This calculation is partially completed
+in Appendix \ref{sec:PXYXYP}.
 
@@
 Here we simplify vectors of the form $H_0\V_M\ket\psi$ to arrive at
-the expansions in \eqref{eq:H_V1_psi_start} and \eqref{eq:H_V2_psi}.
+the expansions in \eqref{eq:H_V1_psi} and \eqref{eq:H_V2_psi}.
 
@@
   = \sum_{\substack{k\\p<q}} h_{pq} v_k \Pi_{pq} X_k \ket\psi,
-  \label{eq:H_V1_psi_start_apndx}
+  \label{eq:H_V1_psi_start}
 \end{align}
@@
 where we have used the facts that $h_{pq}=h_{qp}$ and $h_{pp}=0$.  The
-terms in \eqref{eq:H_V1_psi_start_apndx} with $k\notin\set{p,q}$ are
-then
+terms in \eqref{eq:H_V1_psi_start} with $k\notin\set{p,q}$ are then
 \begin{align}
@@
   = \sum_{k,\ell,q} w_{k\ell} h_{kq} X_q Y_\ell \ket\psi
-  - \sum_{k,\ell} w_{k\ell} h_{k\ell} X_k \odot Y_k \ket\psi
+  - \sum_{k,\ell} w_{k\ell} h_{k\ell} Z_k \ket\psi
 \end{align}
-where $A\odot B \equiv \p{AB+BA}/2$ is a symmetrized product.  In
-total,
+where $Z_k\equiv\p{X_kY_k+Y_kX_k}/2$.  In total,
 \begin{align}
@@
   + \sum_{k,\ell,q} h_{kq} w_{q\ell} X_k Y_\ell \ket\psi
-  - \sum_{k,\ell} w_{k\ell} h_{k\ell} X_k \odot Y_k \ket\psi.
+  - \sum_{k,\ell} w_{k\ell} h_{k\ell} Z_k \ket\psi.
 \end{align}
@@
   \P_0 \V_1\p{\v v}^2 \P_0
-  \FSEQ \sum_{p,q} v_p v_q X_p X_q
+  \EQFS \sum_{p,q} v_p v_q X_p X_q
   = \sum_p v_p^2 X_p^2
@@
 \end{align}
-where $\FSEQ$ denotes equality under a restriction to the fully
+where $\EQFS$ denotes equality under a restriction to the fully
 symmetric manifold $\M_0$; and the coefficients $\v v$ satisfy
@@
   \sum_p v_p^2 X_p^2
-  \FSEQ \sum_p v_p^2 X_0^2
-  \FSEQ \var\p{\v v} \X^{(2)},
-  \label{eq:PXXP_eq}
-\end{align}
-where
-\begin{align}
-  \var\p{\v v} \equiv \EE_p\sp{\p{v_p-\EE_q\sp{v_q}}^2}
-  = \EE_p\sp{v_p^2} - \EE_q\sp{v_q}^2,
+  \EQFS \sum_p v_p^2 X_0^2
+  \EQFS \f{\v v\c\v v}{N} \X^{(2)},
   &&
   \X^{(2)} \equiv \sum_p X_p^2.
+  \label{eq:PXXP_eq}
 \end{align}
@@
   \sum_{p\ne q} v_p v_q X_p X_q
-  \FSEQ \sum_{p\ne q} v_p v_q X_0 X_1,
+  \EQFS \sum_{p\ne q} v_p v_q X_0 X_1,
   \label{eq:PXXP_neq_start}
@@
   = \sum_{p,q} v_p v_q - \sum_{p=q} v_p v_q
-  = - \sum_p v_p^2 = - N \var\p{\v v}.
+  = - \v v\c\v v
 \end{align}
@@
   + \sum_{p\ne q} X_p X_q
-  \FSEQ \X^{(2)} + N \p{N-1} X_0 X_1,
+  \EQFS \X^{(2)} + N \p{N-1} X_0 X_1,
   \label{eq:PXXP_neq_ops}
@@
   \sum_{p\ne q} v_p v_q X_p X_q
-  \FSEQ - N \var\p{\v v} X_0 X_1
-  \FSEQ -\f{\var\p{\v v}}{N-1} \p{\X^2 - \X^{(2)}}
+  \EQFS - \v v\c\v v X_0 X_1
+  \EQFS -\f{\v v\c\v v}{N\p{N-1}} \p{\X^2 - \X^{(2)}}
   \label{eq:PXXP_neq}
@@
 \begin{align}
-  \V_1\p{\v v}^2
-  \FSEQ -\f{\var\p{\v v}}{N-1} \p{\X^2 - N \X^{(2)}}.
+  \V_1\p{\v v}^2 \EQFS -\f{\v v\c\v v}{N\p{N-1}} \p{\X^2 - N \X^{(2)}}.
 \end{align}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Two-body product}
+\subsection{Two-body product, first order}
 \label{sec:PXYP}
@@
   \P_0 \V_2 \P_0
-  \FSEQ \f12 \sum_{p\ne q} w_{pq} X_p Y_q
+  \EQFS \f12 \sum_{p\ne q} w_{pq} X_p Y_q
   = {N\choose 2} \bar w X_0 Y_1,
@@
 \end{align}
-where $\FSEQ$ denotes equality under a restriction to the fully
+where $\EQFS$ denotes equality under a restriction to the fully
 symmetric manifold $\M_0$, ${N\choose2}=N\p{N-1}/2$ is a binomial
@@
   &&
-  \O_{X,Y} \equiv \sum_p X_p \odot Y_p,
+  Z_p \equiv X_p \odot Y_p,
+  &&
+  \Z \equiv \sum_p Z_p,
 \end{align}
@@
 \begin{align}
-  \X\odot\Y \FSEQ \O_{X,Y} + N\p{N-1} X_0 Y_1.
+  \X\odot\Y \EQFS \Z + N\p{N-1} X_0 Y_1.
 \end{align}
@@
 \begin{align}
-  \V_2 \FSEQ \f12 \bar w \p{\X\odot\Y -  \O_{X,Y}}.
+  \V_2 \EQFS \f12 \bar w \p{\X\odot\Y -  \Z}.
 \end{align}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Two-body product, second order}
+\label{sec:PXYXYP}
+
+Calculating the second-order effective Hamiltonian $H_2^{(2)}$ induced
+on the fully symmetric manifold $\M_0$ by the perturbation $\V_2$ in
+\eqref{eq:perturbations} requires us to simplify
+\begin{align}
+  \P_0 \V_2^2 \P_0
+  \EQFS \f14 \sum_{k,\ell,p,q}
+  w_{k\ell} w_{pq} X_k Y_\ell X_p Y_q,
+  \label{eq:PXYXYP_start}
+\end{align}
+where $\EQFS$ denotes equality under a restriction to the fully
+symmetric manifold $\M_0$, and the coefficients $\m w$ satisfy
+\eqref{eq:cond_2} with eigenvalue $\Delta\p{\m w}\ne0$, which implies
+that $\m w$ is mean-zero.  To simplify the sum in
+\eqref{eq:PXYXYP_start}, we first consider the terms with
+$p,q\in\set{k,\ell}$:
+\begin{align}
+  \sum_{\substack{k,\ell\\p,q\in\set{k,\ell}}}
+  w_{k\ell} w_{pq} X_k Y_\ell X_p Y_q
+  \EQFS \v{\m w}\c\v{\m w} \p{X_0^2 Y_1^2 + X_0 Y_0 Y_1 X_1},
+  &&
+  \v{\m w} \equiv \sum_{k,\ell} w_{k\ell} \ket{k\ell}.
+\end{align}
+The terms of the sum in \eqref{eq:PXYXYP_start} with
+$p\in\set{k,\ell}$ and $q\notin\set{k,\ell}$ or vice versa are
+\begin{align}
+  \sum_{\substack{k,\ell\\p\in\set{k,\ell}\\q\notin\set{k,\ell}}}
+  w_{k\ell} w_{pq} X_k Y_\ell X_p Y_q
+  &\EQFS \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}}
+  w_{k\ell} w_{k q} \p{X_0^2 Y_1 Y_2 + Y_0 X_0 X_1 Y_2}, \\
+  \sum_{\substack{k,\ell\\p\notin\set{k,\ell}\\q\in\set{k,\ell}}}
+  w_{k\ell} w_{pq} X_k Y_\ell X_p Y_q
+  &\EQFS \sum_{\substack{k,\ell\\p\notin\set{k,\ell}}}
+  w_{k\ell} w_{pk} \p{X_0 Y_0 X_1 Y_2 + Y_0^2 X_1 X_2},
+\end{align}
+where
+\begin{align}
+  \sum_{\substack{k,\ell\\p\notin\set{k,\ell}}} w_{k\ell} w_{pk}
+  = \sum_{k,\ell,p} w_{k\ell} w_{kp} - \sum_{k,\ell} w_{k\ell}^2
+  = \v w\c\v w - \v{\m w}\c\v{\m w},
+  &&
+  \v w \equiv \sum_{k,\ell} w_{k\ell} \ket{k}.
+\end{align}
+Finally, the terms of the sum in \eqref{eq:PXYXYP_start} with
+$p,q\notin\set{k,\ell}$ are
+\begin{align}
+  \sum_{\substack{k,\ell\\p,q\notin\set{k,\ell}}}
+  w_{k\ell} w_{pq} X_k Y_\ell X_p Y_q
+  \EQFS \sum_{\substack{k,\ell\\p,q\notin\set{k,\ell}}}
+  w_{k\ell} w_{pq} X_0 X_1 Y_2 Y_3,
+\end{align}
+where, using the fact that $\m w$ is mean-zero,
+\begin{align}
+  \sum_{\substack{k,\ell\\p,q\notin\set{k,\ell}}} w_{k\ell} w_{pq}
+  = - \sum_{\substack{k,\ell\\p,q\in\set{k,\ell}}} w_{k\ell} w_{pq}
+  - \sum_{\substack{k,\ell\\p\in\set{k,\ell}\\q\notin\set{k,\ell}}}
+  w_{k\ell} w_{pq}
+  - \sum_{\substack{k,\ell\\p\notin\set{k,\ell}\\q\in\set{k,\ell}}}
+  w_{k\ell} w_{pq}
+  = 2 \v{\m w}\c\v{\m w} - 4 \v w\c\v w.
+\end{align}
+Altogether, we thus far have that
+\begin{multline}
+  \P_0 \V_2^2 \P_0
+  \EQFS \p{\f12 \v{\m w}\c\v{\m w} - \v w\c\v w} X_0 X_1 Y_2 Y_3
+  + \f14 \v{\m w}\c\v{\m w} \p{X_0^2 Y_1^2 + X_0 Y_0 Y_1 X_1} \\
+  + \f14 \p{\v w\c\v w - \v{\m w}\c\v{\m w}}
+  \p{X_0^2 Y_1 Y_2 + X_0 Y_0 X_1 Y_2 + Y_0 X_0 X_1 Y_2 + Y_0^2 X_1 X_2}.
+\end{multline}
 

### 2019-11-13 collective_sun/sun_perturbation.tex
changed notation for single-spin vs. collective operators the new notation should play more nicely with the second-order effect of two-body perturbations

@@
 \newcommand{\EQFS}{=_{\text{FS}}}
+\newcommand{\col}{\underline}
 
@@
 \end{align}
-where $X_p,Y_p$ are trace-zero operators on spin $p$, with e.g.~$X_p$
-related to $X_q$ by a permutation of spins $p$ and $q$; $\v X,\v Y$
+where $X,Y$ are trace-zero operators on an $n$-dimensional Hilbert
+space; $X_p,Y_p$ denote the action of $X,Y$ on spin $p$; $\v X,\v Y$
 are respectively vectors of all $X_p,Y_p$; $\v v$ and $\m w$ are
@@
   H_1^{(1)} = \sum_p v_p \P_0 X_p \P_0
-  = \sum_p v_p \P_0 X_0 \P_0 \EQFS \bar v \X,
+  = \sum_p v_p \P_0 X_0 \P_0 \EQFS \bar v \col{X},
   \label{eq:H_1_1}
@@
   &&
-  \X \equiv \sum_p X_p.
+  \col{X} \equiv \sum_p X_p.
 \end{align}
@@
 $\v v_\Delta$, and is entirely independent of the state $\ket\psi$ or
-choice of single-spin trace-zero operator $X_p$ used to build $\V_1$
-in \eqref{eq:perturbations}.  Finding operators $\V_1\p{\v v_\Delta}$
+choice of single-spin trace-zero operator $X$ used to build $\V_1$ in
+\eqref{eq:perturbations}.  Finding operators $\V_1\p{\v v_\Delta}$
 that generate eigenvectors of the interaction Hamiltonian $H_0$ when
@@
 \begin{align}
-  H_1^{(2)} \EQFS \tilde v \p{\X^2 - N \X^{(2)}},
+  H_1^{(2)} \EQFS \tilde v \sp{\p{\col{X}}^2 - N \p{\col{X^2}}},
   \label{eq:H_1_2}
@@
   &&
-  \X^{(2)} \equiv \sum_p X_p^2.
+  \col{X^2} \equiv \sum_p X_p^2.
 \end{align}
@@
   = \P_0 \V_2 \P_0
-  \EQFS \f12 \bar w \p{\X\odot\Y - \Z},
+  \EQFS \f12 \bar w \p{\col{X}\odot\col{Y} - \col{X\odot Y}},
 \end{align}
@@
   &&
-  \Z \equiv \sum_p Z_p,
+  \col{X} \equiv \sum_p X_p,
   &&
-  Z_p \equiv X_p \odot Y_p.
+  \col{Y} \equiv \sum_p Y_p,
+  &&
+  \col{X\odot Y} \equiv \sum_p X_p\odot Y_p.
 \end{align}
@@
   + \sum_{k,\ell,q} h_{kq} w_{q\ell} X_k Y_\ell \ket\psi
-  - \sum_{k,\ell} w_{k\ell} h_{k\ell} Z_k \ket\psi \\
+  - \sum_{k,\ell} w_{k\ell} h_{k\ell} X_k \odot Y_k \ket\psi \\
   &= E_0 \V_2 \ket\psi
   + \v X \c \p{\bar{\m h}\oc\m w + \m h\c\m w} \c \v Y \ket\psi
-  - \v\1 \c \p{\m h\oc\m w} \c \v Z \ket\psi,
+  - \v\1 \c \p{\m h\oc\m w} \c \v{X\odot Y} \ket\psi,
   \label{eq:H_V2_psi}
 \end{align}
-where $\m w,\m h,\bar{\m h}$ are matrices of all
-$w_{k\ell},h_{k\ell},\bar h_{k\ell}$, with
+where to keep notation compact, we define
 \begin{align}
@@
 \end{align}
-$\m x\oc\m y$ is an element-wise (Hadamard) product of $\m x$ and
-$\m y$, i.e.~with matrix elements $x_{k\ell} y_{k\ell}$;
-$\v\1\equiv\p{1,1,1,\cdots}$ is an $N$-component vector of ones; and
-$\v Z$ is a vector of all $Z_k$.  We thus find that the vector
-$\V_2\ket\psi$ is an eigenvector of the interaction Hamiltonian $H_0$
-if the coefficient matrix $\m w$ of all $w_{pq}$ satisfies
+$\m w,\m h,\bar{\m h}$ are matrices of all
+$w_{k\ell},h_{k\ell},\bar h_{k\ell}$; $\m x\oc\m y$ is an element-wise
+(Hadamard) product of $\m x$ and $\m y$, i.e.~with matrix elements
+$x_{k\ell} y_{k\ell}$; $\v\1\equiv\p{1,1,1,\cdots}$ is an
+$N$-component vector of ones; and $\v{X\odot Y}$ is a vector of all
+$X_k\odot Y_k$.  We thus find that the vector $\V_2\ket\psi$ is an
+eigenvector of the interaction Hamiltonian $H_0$ if the coefficient
+matrix $\m w$ of all $w_{pq}$ satisfies
 \begin{align}
@@
 components of the condition in \eqref{eq:cond_2} automatically enforce
-that the last term ($\sim\v Z$) in \eqref{eq:H_V2_psi} vanishes.
+that the last term ($\sim\v{X\odot Y}$) in \eqref{eq:H_V2_psi}
+vanishes.
 
@@
   = \sum_{k,\ell,q} w_{k\ell} h_{kq} X_q Y_\ell \ket\psi
-  - \sum_{k,\ell} w_{k\ell} h_{k\ell} Z_k \ket\psi
+  - \sum_{k,\ell} w_{k\ell} h_{k\ell} X_k\odot Y_k \ket\psi
 \end{align}
@@
   + \sum_{k,\ell,q} h_{kq} w_{q\ell} X_k Y_\ell \ket\psi
-  - \sum_{k,\ell} w_{k\ell} h_{k\ell} Z_k \ket\psi.
+  - \sum_{k,\ell} w_{k\ell} h_{k\ell} X_k\odot Y_k \ket\psi.
 \end{align}
@@
   \EQFS \sum_p v_p^2 X_0^2
-  \EQFS \f{\v v\c\v v}{N} \X^{(2)},
+  \EQFS \f1N \v v\c\v v\, \col{X^2},
   &&
-  \X^{(2)} \equiv \sum_p X_p^2.
+  \col{X^2} \equiv \sum_p X_p^2.
   \label{eq:PXXP_eq}
@@
 \begin{align}
-  \X^2
-  = \sum_p X_p^2
-  + \sum_{p\ne q} X_p X_q
-  \EQFS \X^{(2)} + N \p{N-1} X_0 X_1,
+  \p{\col{X}}^2
+  = \sum_p X_p^2 + \sum_{p\ne q} X_p X_q
+  \EQFS \col{X^2} + N \p{N-1} X_0 X_1,
   \label{eq:PXXP_neq_ops}
@@
   \EQFS - \v v\c\v v X_0 X_1
-  \EQFS -\f{\v v\c\v v}{N\p{N-1}} \p{\X^2 - \X^{(2)}}
+  \EQFS -\f{\v v\c\v v}{N\p{N-1}}
+  \sp{\p{\col{X}}^2 - N \p{\col{X^2}}}
   \label{eq:PXXP_neq}
@@
 \begin{align}
-  \V_1\p{\v v}^2 \EQFS -\f{\v v\c\v v}{N\p{N-1}} \p{\X^2 - N \X^{(2)}}.
+  \V_1\p{\v v}^2
+  \EQFS -\f{\v v\c\v v}{N\p{N-1}} \sp{\p{\col{X}}^2 - N \p{\col{X^2}}}.
 \end{align}
@@
 \begin{align}
-  \X \equiv \sum_p X_p,
+  \col{X} \equiv \sum_p X_p,
   &&
-  \Y \equiv \sum_p Y_p,
+  \col{Y} \equiv \sum_p Y_p,
   &&
@@
   &&
-  Z_p \equiv X_p \odot Y_p,
-  &&
-  \Z \equiv \sum_p Z_p,
+  \col{X\odot Y} \equiv \sum_p X_p \odot Y_p,
 \end{align}
@@
 \begin{align}
-  \X\odot\Y \EQFS \Z + N\p{N-1} X_0 Y_1.
+  \col{X}\odot\col{Y} \EQFS \col{X\odot Y} + N\p{N-1} X_0 Y_1.
 \end{align}
@@
 \begin{align}
-  \V_2 \EQFS \f12 \bar w \p{\X\odot\Y -  \Z}.
+  \V_2 \EQFS \f12 \bar w \p{\col{X}\odot\col{Y} - \col{X\odot Y}}.
 \end{align}
@@
   \P_0 \V_2^2 \P_0
-  \EQFS \p{\f12 \v{\m w}\c\v{\m w} - \v w\c\v w} X_0 X_1 Y_2 Y_3
+  \EQFS -\p{\v w\c\v w - \f12 \v{\m w}\c\v{\m w}} X_0 X_1 Y_2 Y_3
   + \f14 \v{\m w}\c\v{\m w} \p{X_0^2 Y_1^2 + X_0 Y_0 Y_1 X_1} \\
@@
 \end{multline}
+If $X=Y=s_\z$, where $s_\z\equiv\p{\op\up-\op\dn}/2$ is an SU(2) spin
+operator, then
+\begin{align}
+  \P_0 \V_2^2 \P_0
+  \EQFS - \sp{\f{\v w\c\v w-\v{\m w}\c\v{\m w}/2}{N_4}} S_\z^4
+  + \sp{\f{\v w\c\v w-\v{\m w}\c\v{\m w}}{4N_2}
+    + \f{\p{3N-4}\p{\v w\c\v w - \v{\m w}\c\v{\m w}/2}}{2N_4}} S_\z^2,
+\end{align}
+where $S_\z\equiv\sum_p s_\z^{(p)}$ is a collective SU(2) spin
+operator,
+\begin{align}
+  N_k \equiv \prod_{j=0}^{k-1} \p{N-j} = \f{N!}{\p{N-k}!},
+\end{align}
+and we have neglected scalar terms with no physical consequence.
 

### 2019-11-15 collective_sun/sun_perturbation.tex
added second-order results for ZZ interactions, and some second-order results for a general two-body perturbation

@@
 \begin{align}
-  H_1^{(2)} \EQFS \tilde v \sp{\p{\col{X}}^2 - N \p{\col{X^2}}},
+  H_1^{(2)}
+  \EQFS \f1{N\p{N-1}}
+  \sum_{\Delta>0} \f{\v v_\Delta\c\v v_\Delta}{\Delta}
+  \sp{\p{\col{X}}^2 - N \p{\col{X^2}}},
   \label{eq:H_1_2}
@@
 where $\EQFS$ denotes equality under a restriction to the fully
-symmetric manifold $\M_0$, and
+symmetric manifold $\M_0$, $\v v_\Delta$ is the projection of $\v v$
+onto the $\Delta$-eigenspace of the matrix $\m h-\diag\v h$, and
 \begin{align}
-  \tilde v \equiv \f1{N\p{N-1}}
-  \sum_{\Delta>0} \f{\v v_\Delta\c\v v_\Delta}{\Delta},
+  \col{X} \equiv \sum_p X_p,
   &&
@@
 \end{align}
-As the vectors $\v v_\Delta$ with $\Delta>0$ are mean-zero, we can
-equivalently write
-\begin{align}
-  \tilde v = \f1{N-1}
-  \sum_{\Delta>0} \f{\var\p{\v v_\Delta}}{\Delta},
-\end{align}
-where $\var\p{\v x}$ is the variance of the entries in $\v x$.
 
@@
 If we {\it construct} a perturbation $\V_2\p{\m w_\Delta}$ with
-coefficients $\m w_\Delta$ that satisfy \eqref{eq:cond_2} or
-\eqref{eq:cond_2_eig} with eigenvalue $\Delta/2$, then
+coefficients $\m w_\Delta$ that satisfy \eqref{eq:cond_2_eig} with
+eigenvalue $\Delta/2$, then
 \begin{align}
@@
 matrix $\m w$ into its projections $\m w_\Delta$ onto the eigenspace
-of matrices satisfying \eqref{eq:cond_2} with eigenvalue $\Delta/2$,
-i.e.~$\m w=\sum_\Delta\m w_\Delta$, which allows us to expand
+of matrices satisfying \eqref{eq:cond_2_eig} with eigenvalue
+$\Delta/2$, i.e.~$\m w=\sum_\Delta\m w_\Delta$, which allows us to
+expand
 \begin{align}
@@
 \end{align}
-and in turn simplify
+and in turn write
 \begin{align}
@@
 \end{align}
-where the product $\P_0 \V_2\p{\m w_\Delta}^2 \P_0$ can be written in
-terms collective operators using techniques similar to those in
-Appendix \ref{sec:sym_prod}.  This calculation is partially completed
-in Appendix \ref{sec:PXYXYP}.
+where the product $\P_0 \V_2\p{\m w_\Delta}^2 \P_0$ is simplified in
+Appendix \ref{sec:PXYXYP}.  If $X=Y=s_\z$, where
+$s_\z\equiv\p{\op\up-\op\dn}/2$ is an SU(2) spin operator, then we can expand
+\begin{align}
+  H_2^{(2)} \EQFS W_4\p{\m h,\m w} S_\z^4 - W_2\p{\m h,\m w} S_\z^2,
+\end{align}
+where $S_\z\equiv\sum_p s_\z^{(p)}$ is a collective SU(2) spin
+operator, $W_k\p{\m h,\m w}$ are scalar coefficients, and we have
+neglected constant terms in $H_2^{(2)}$ without physical consequence.
+In order to expand the coefficients $W_k\p{\m h,\m w}$, for any matrix
+$\m m$ we define the vectors
+\begin{align}
+  \v{\m m} \equiv \sum_{p,q} m_{pq} \ket{pq},
+  &&
+  \v m \equiv \sum_{p,q} m_{pq} \ket{p},
+\end{align}
+using which we have
+\begin{align}
+  W_4\p{\m h,\m w}
+  \equiv \sum_\Delta \f1\Delta \f{\v w_\Delta\c\v w_\Delta
+    - \v{\m w}_\Delta\c\v{\m w}_\Delta/2}{N!_4},
+\end{align}
+\begin{align}
+  W_2\p{\m h,\m w}
+  \equiv \sum_\Delta \f1{\Delta} \sp{\f{\v w_\Delta\c\v w_\Delta
+      - \v{\m w}_\Delta\c\v{\m w}_\Delta}{4N!_2}
+    + \f{\p{3N-4}\p{\v w_\Delta\c\v w_\Delta
+        - \v{\m w}_\Delta\c\v{\m w}_\Delta/2}}{2N!_4}},
+\end{align}
+where
+\begin{align}
+  N!_k \equiv \prod_{j=0}^{k-1} \p{N-j} = \f{N!}{\p{N-k}!}.
+\end{align}
+Note that the dependence of the coefficients of $W_k\p{\m h,\m w}$ on
+the interaction matrix $\m h$ is implicit in the decomposition of the
+coefficient matrix $\m w$ into components $\m w_\Delta$ as
+$\m w=\sum_\Delta\m w_\Delta$, where $\m w_\Delta$ is the projection
+of $\m w$ onto the $\p{\Delta/2}$-eigenspace of the operator
+$\tilde{\m h}$ defined in \eqref{eq:h_super_mat}.
 
@@
 \begin{align}
-  \V_1\p{\v v}^2
-  \EQFS -\f{\v v\c\v v}{N\p{N-1}} \sp{\p{\col{X}}^2 - N \p{\col{X^2}}}.
+  \P_0 \V_1\p{\v v}^2 \P_0
+  \EQFS -\f{\v v\c\v v}{N\p{N-1}}
+  \sp{\p{\col{X}}^2 - N \p{\col{X^2}}}.
 \end{align}
@@
 \begin{align}
-  \V_2 \EQFS \f12 \bar w \p{\col{X}\odot\col{Y} - \col{X\odot Y}}.
+  \P_0 \V_2 \P_0
+  \EQFS \f12 \bar w \p{\col{X}\odot\col{Y} - \col{X\odot Y}}.
 \end{align}
@@
 \end{align}
-Altogether, we thus far have that
+Altogether, we thus have that
 \begin{multline}
@@
   + \f14 \p{\v w\c\v w - \v{\m w}\c\v{\m w}}
-  \p{X_0^2 Y_1 Y_2 + X_0 Y_0 X_1 Y_2 + Y_0 X_0 X_1 Y_2 + Y_0^2 X_1 X_2}.
+  \p{X_0^2 Y_1 Y_2 + X_0 Y_0 X_1 Y_2
+    + Y_0 X_0 X_1 Y_2 + Y_0^2 X_1 X_2}.
+  \label{eq:PXYXYP_multi_body}
 \end{multline}
-If $X=Y=s_\z$, where $s_\z\equiv\p{\op\up-\op\dn}/2$ is an SU(2) spin
-operator, then
+Denoting the collective version of any single-spin operator $Z$ by
+$\col{Z}\equiv\sum_pZ_p$, defining the falling factorial
 \begin{align}
-  \P_0 \V_2^2 \P_0
-  \EQFS - \sp{\f{\v w\c\v w-\v{\m w}\c\v{\m w}/2}{N_4}} S_\z^4
-  + \sp{\f{\v w\c\v w-\v{\m w}\c\v{\m w}}{4N_2}
-    + \f{\p{3N-4}\p{\v w\c\v w - \v{\m w}\c\v{\m w}/2}}{2N_4}} S_\z^2,
+  N!_k \equiv \prod_{j=0}^{k-1} \p{N-j} = \f{N!}{\p{N-k}!},
 \end{align}
-where $S_\z\equiv\sum_p s_\z^{(p)}$ is a collective SU(2) spin
-operator,
+denoting symmetrized products by $A\odot B\equiv\p{AB+BA}/2$, and
+adopting the convention that exponents are evaluated before
+symmetrized products,
+i.e.~$A^k\odot B\equiv\p{A^k}\odot B\ne A^{k-1}\p{A\odot B}$, the
+operators in \eqref{eq:PXYXYP_multi_body} can be written in terms of
+collective operators within the fully symmetric manifold through the
+identities
 \begin{align}
-  N_k \equiv \prod_{j=0}^{k-1} \p{N-j} = \f{N!}{\p{N-k}!},
+  N!_2 \p{X_0^2 Y_1^2 + X_0 Y_0 Y_1 X_1}
+  \EQFS \col{X^2}\odot\col{Y^2} + \col{XY} \odot \col{YX}
+  - \col{X^2\odot Y^2} - \col{\p{XY}\odot\p{YX}}
 \end{align}
-and we have neglected scalar terms with no physical consequence.
+\begin{align}
+  N!_3 &\p{X_0^2 Y_1 Y_2 + X_0 Y_0 X_1 Y_2
+    + Y_0 X_0 X_1 Y_2 + Y_0^2 X_1 X_2}
+  \notag \\
+  &\EQFS 4\col{\p{X\odot Y}^2} + 5 \col{X^2\odot Y^2} - \f12\col{\p{XY}\odot\p{YX}}
+  \notag \\
+  &\quad- \f32 \col{X}\odot\col{Y^2X} - \f32 \col{XY^2}\odot\col{X}
+  - \f32 \col{Y}\odot\col{X^2Y} - \f32 \col{YX^2}\odot\col{X}
+  \notag \\
+  &\quad- \col{X}\odot\col{YXY} - \col{Y}\odot\col{XYX}
+  - 2 \col{XY}\odot\col{YX} - 2\col{X^2}\odot\col{Y^2}
+  \notag \\
+  &\quad+ \col{X}\,\col{Y^2}\,\col{X} + \col{Y}\,\col{X^2}\,\col{Y}
+  +\col{XY}\odot\p{\col{Y}\,\col{X}} + \col{YX}\odot\p{\col{X}\,\col{Y}}.
+\end{align}
+
+[UNFINISHED]
 

### 2019-11-19 collective_sun/sun_perturbation.tex
added appendix on converting muti-local operators to collective operators within the fully symmetric manifold

--- 2019-11-13 collective_sun/sun_perturbation.tex
@@
 \newcommand{\I}{\mathcal{I}}
+\renewcommand{\L}{\mathcal{L}}
 \newcommand{\M}{\mathcal{M}}
@@
 
+\newcommand{\PP}{\mathbb{Z}}
+
 \newcommand{\FS}{\text{FS}}
@@
 \DeclareMathOperator{\diag}{diag}
+\DeclareMathOperator{\counts}{counts}
 
@@
 \end{align}
-where the product $\P_0 \V_2\p{\m w_\Delta}^2 \P_0$ can be written in
-terms collective operators using techniques similar to those in
-Appendix \ref{sec:sym_prod}.  This calculation is partially completed
-in Appendix \ref{sec:PXYXYP}.
+where the product $\P_0 \V_2\p{\m w_\Delta}^2 \P_0$ is simplified in
+Appendices \ref{sec:PXYXYP} and \ref{sec:multi_to_collective}.  If
+$X=Y=s_\z$, where $s_\z\equiv\p{\op\up-\op\dn}/2$ is an SU(2) spin
+operator, then we can expand
+\begin{align}
+  H_2^{(2)} \EQFS W_4\p{\m h,\m w} S_\z^4 - W_2\p{\m h,\m w} S_\z^2,
+\end{align}
+where $S_\z\equiv\sum_p s_\z^{(p)}$ is a collective SU(2) spin
+operator, $W_k\p{\m h,\m w}$ are scalar coefficients, and we have
+neglected scalar terms in $H_2^{(2)}$ without physical consequence.
+In order to expand the coefficients $W_k\p{\m h,\m w}$, for any matrix
+$\m m$ we define the vectors
+\begin{align}
+  \v{\m m} \equiv \sum_{p,q} m_{pq} \ket{pq},
+  &&
+  \v m \equiv \sum_{p,q} m_{pq} \ket{p},
+\end{align}
+using which we have
+\begin{align}
+  W_4\p{\m h,\m w}
+  \equiv \sum_\Delta \f1\Delta \f{\v w_\Delta\c\v w_\Delta
+    - \v{\m w}_\Delta\c\v{\m w}_\Delta/2}{N!_4},
+\end{align}
+\begin{align}
+  W_2\p{\m h,\m w}
+  \equiv \sum_\Delta \f1{\Delta} \sp{\f{\v w_\Delta\c\v w_\Delta
+      - \v{\m w}_\Delta\c\v{\m w}_\Delta}{4N!_2}
+    + \f{\p{3N-4}\p{\v w_\Delta\c\v w_\Delta
+        - \v{\m w}_\Delta\c\v{\m w}_\Delta/2}}{2N!_4}},
+\end{align}
+where
+\begin{align}
+  N!_k \equiv \prod_{j=0}^{k-1} \p{N-j} = \f{N!}{\p{N-k}!}.
+\end{align}
+Note that the dependence of the coefficients of $W_k\p{\m h,\m w}$ on
+the interaction matrix $\m h$ is implicit in the decomposition of the
+coefficient matrix $\m w$ into components $\m w_\Delta$ as
+$\m w=\sum_\Delta\m w_\Delta$, where $\m w_\Delta$ is the projection
+of $\m w$ onto the $\p{\Delta/2}$-eigenspace of the operator
+$\tilde{\m h}$ defined in \eqref{eq:h_super_mat}.
 
@@
   + \f14 \p{\v w\c\v w - \v{\m w}\c\v{\m w}}
-  \p{X_0^2 Y_1 Y_2 + X_0 Y_0 X_1 Y_2 + Y_0 X_0 X_1 Y_2 + Y_0^2 X_1 X_2}.
+  \p{X_0^2 Y_1 Y_2 + X_0 Y_0 X_1 Y_2
+    + Y_0 X_0 X_1 Y_2 + Y_0^2 X_1 X_2}.
+  \label{eq:PXYXYP}
 \end{multline}
-If $X=Y=s_\z$, where $s_\z\equiv\p{\op\up-\op\dn}/2$ is an SU(2) spin
-operator, then
+In principle, our next task is to write this result in terms of
+collective operators within the fully symmetric manifold; this task
+turns out to be rather non-trivial in general.  We therefore discuss
+the decomposition of multi-local operators (e.g.~$X_0Y_0X_1Y_2$, a
+3-local operator) into a sum of collective operators within the fully
+symmetric manifold in Appendix \ref{sec:multi_to_collective}.  At the
+end of Appendix \ref{sec:multi_to_collective}, we provide analytical
+expressions for the special cases of
+\begin{enumerate*}
+\item multi-local operators that are tensor powers of a single
+  operator, i.e.~of the form $X^{\otimes k}$; and
+\item multi-local operators of the form $X^{\otimes k}$ with $X^2=1$.
+\end{enumerate*}
+These cases are relevant for e.g.~the SU(2) case with Ising-type
+interactions of the form $Z\otimes Z$, with $Z$ a single-spin Pauli
+operator.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Collective and multi-local operators in the fully symmetric
+  manifold}
+\label{sec:multi_to_collective}
+
+Here we provide a decomposition multi-local operators into sums of
+collective operators (and vice versa) within the fully symmetric
+manifold of $N$ spins.  Multi-local operators are operators of the
+form
 \begin{align}
-  \P_0 \V_2^2 \P_0
-  \EQFS - \sp{\f{\v w\c\v w-\v{\m w}\c\v{\m w}/2}{N_4}} S_\z^4
-  + \sp{\f{\v w\c\v w-\v{\m w}\c\v{\m w}}{4N_2}
-    + \f{\p{3N-4}\p{\v w\c\v w - \v{\m w}\c\v{\m w}/2}}{2N_4}} S_\z^2,
+  O_S \equiv \bigotimes_{O\in S} O,
+  \label{eq:multi_op_set}
 \end{align}
-where $S_\z\equiv\sum_p s_\z^{(p)}$ is a collective SU(2) spin
-operator,
+where $S\equiv\set{O}$ is a set of single-spin operators $O$.  The
+order of tensor factors in \eqref{eq:multi_op_set} does not matter
+after projecting operators onto the fully symmetric manifold, as we
+always will in this section.  While the decomposition of a collective
+operator into a sum of multi-local operators is unique, the opposite
+is not true: there is generally no unique decomposition of a
+multi-local operator into a sum of collective operators.  We can,
+however, enforce a canonical decomposition by choosing a set of
+collective operators that obey the same symmetries as $O_S$, namely
+invariance under arbitrary permutations of the single-spin operators
+$O\in S$.  To this end, for any partition\footnote{A partition
+  $P\equiv\set{p}$ of $S$ consists of mutually disjoint, non-empty
+  subsets $p\subset S$ with $\bigcup_{p\in P}p=S$.} $P\equiv\set{p}$
+of $S$ we define the $\abs{P}$-local operator
 \begin{align}
-  N_k \equiv \prod_{j=0}^{k-1} \p{N-j} = \f{N!}{\p{N-k}!},
+  O_P \equiv \bigotimes_{p\in P} \bigodot_{O\in p} O,
 \end{align}
-and we have neglected scalar terms with no physical consequence.
+where $\bigodot_{x\in\X} x$ denotes a symmetrized product of all
+elements in $\X$, i.e.
+\begin{align}
+  \bigodot_{x\in\X} x
+  \equiv \f1{\abs{\X}!}
+  \sum_{\substack{\t{permutations}\\\sigma~\t{of}~\X}}
+  ~ \prod_{k=1}^{\abs{\X}} x_{\sigma_k},
+\end{align}
+with $\sigma_k$ the $k$-th element of the permutation $\sigma$ when
+expressed in one-line notation; for example,
+\begin{align}
+  \bigodot_{j=1}^2 x_j &= x_1 \odot x_2 = \f12 \p{x_1x_2 + x_2x_1}, \\
+  \bigodot_{j=1}^3 x_j
+  &= \f1{3!} \p{x_1x_2x_3 + x_1x_3x_2 + x_2x_1x_3
+    + x_2x_3x_1 + x_3x_1x_2 + x_3x_2x_1}.
+\end{align}
+The set of multi-local operators $\set{O_P:\t{partitions}~P~\t{of}~S}$
+consists of all operators that
+\begin{enumerate*}
+\item can be built from ordinary and tensor products of all
+  single-spin operators $O\in S$, and
+\item are invariant under arbitrary permutations of $S$ (after
+  projection onto the fully symmetric manifold).
+\end{enumerate*}
+For any set $S$ of single-spin operators or partition $P$ of $S$, we
+can similarly define the collective operators
+\begin{align}
+  \O_S \equiv \bigodot_{O\in S} \col{O},
+  &&
+  \O_P \equiv \bigodot_{p\in P} \col{\textstyle\bigodot_{O\in p} O},
+\end{align}
+where we underline a single-spin operator to denote its collective
+version, i.e.~$\col{Q}\equiv\sum_j Q^{(j)}$, where $Q^{(j)}$ is an
+operator that acts with the single-spin operator $Q$ on spin $j$ and
+trivially on all other spins.  By construction, the set of collective
+operators $\set{\O_P:\t{partitions}~P~\t{of}~S}$ is one-to-one to the
+set of multi-local operators $\set{O_P:\t{partitions}~P~\t{of}~S}$,
+and it is straightforward to work out that for any set $S$ of
+single-spin operators,
+\begin{align}
+  \O_S
+  \EQFS \sum_{\substack{\t{partitions}\\P~\t{of}~S}}
+  N!_{\abs{P}}\, O_P,
+  &&
+  N!_k \equiv \prod_{j=0}^{k-1}\p{N-j}! = \f{N!}{\p{N-k}!},
+  \label{eq:collective_to_multi_full}
+\end{align}
+where $\EQFS$ denotes equality under a restriction to the fully
+symmetric manifold.  For any multi-local operator $O_S$ indexed by a
+set $S$ of single-spin operators, we can collect expansions of the
+form \eqref{eq:collective_to_multi_full} for all collective operators
+built from non-empty subsets of $S$ into a system of linear equations
+$\L$, and in turn solve $\L$ for $O_S$.  When
+$\abs{S}\in\set{1,2,3,4}$ (and, by conjecture, all $S$), by use of a
+computer algebra system (namely, Mathematica) we find that
+\begin{align}
+  N!_{\abs{S}}\, O_S
+  &\EQFS \sum_{\substack{\t{partitions}\\P~\t{of}~S}}
+  \p{-1}^{\abs{S}-\abs{P}} \O_P \times \prod_{p\in P}\p{\abs{p}-1}!
+  \label{eq:multi_to_collective_full} \\
+  &\EQFS \f1{\abs{S}!} \sum_{\substack{\t{permutations}\\\Pi~\t{of}~S}}
+  \sum_{\substack{\t{ordered}\\\t{partitions}\\\tilde P~\t{of}~\Pi}}
+  \p{-1}^{\abs{S}-\abs{\tilde P}} \O_{\tilde P}
+  \times \prod_{p\in\tilde P} \p{\abs{p}-1}!
+  \label{eq:multi_to_collective_full_simp}
+\end{align}
+where we identify permutations $\Pi$ of $S$ with an ordered list
+(tuple) $\p{O_1,O_2,\cdots,O_{\abs{S}}}$ of all single-spin operators
+in $S$, and we enforce that partitions $\tilde P$ of an ordered list
+$\Pi$, as well as parts $p\in\tilde P$, are lexicographically ordered
+in accordance with the ordering of operators in $\Pi$.  If
+$S=\set{X,Y,Z}$, for example, then the ordered partitions of the
+trivial permutation $\Pi=\p{X,Y,Z}$ are
+\begin{align}
+  \p{\p{X,Y,Z}}, &&
+  \p{\p{X,Y},\p{Z}}, &&
+  \p{\p{X,Z},\p{Y}}, &&
+  \p{\p{X},\p{Y,Z}}, &&
+  \p{\p{X},\p{Y},\p{Z}}.
+\end{align}
+For ordered lists $\tilde S$ and ordered partitions $\tilde P$, we
+define
+\begin{align}
+  O_{\tilde S} \equiv \bigotimes_{O\in\tilde S} O,
+  &&
+  O_{\tilde P} \equiv \bigotimes_{p\in\tilde P} \prod_{O\in p} O,
+  &&
+  \O_{\tilde S} \equiv \prod_{O\in\tilde S} \col{O},
+  &&
+  \O_{\tilde P}
+  \equiv \prod_{p\in\tilde P} \col{\textstyle\prod_{O\in p} O},
+\end{align}
+in which case we can also expand
+\begin{align}
+  \O_{\tilde S} \EQFS
+  \sum_{\substack{\t{ordered}\\\t{partitions}\\\tilde P~\t{of}~\tilde S}}
+  N!_{\abs{\tilde P}} O_{\tilde P},
+\end{align}
+in analogy to \eqref{eq:collective_to_multi_full}.
+
+While the expansions in \eqref{eq:multi_to_collective_full} and
+\eqref{eq:multi_to_collective_full_simp} closely resemble each other,
+in practice the latter involves less computational overhead due to the
+lack of symmetrization in the construction of $\O_{\tilde P}$.  There
+are, in principle, even more economical decompositions of a
+multi-local operators into collective opreators, for example
+\begin{align}
+  N!_2\, X\otimes Y
+  &\EQFS \col{X}\,\col{Y} - \col{XY}, \\
+  N!_3\, X\otimes Y \otimes Z
+  &\EQFS \col{X}\,\col{Y}\,\col{Z}
+  - \col{X}\,\col{YZ} - \col{Y}\,\col{XZ} - \col{Z}\,\col{XY}
+  + \col{ZXY} + \col{YXZ},
+\end{align}
+but these decompositions generally do not preserve the manifest
+symmetry of a multi-local operator $O_S$ under arbitrary permutations
+of $S$, and we do not know of any general prescription to arrive at
+such decompositions aside from solving systems of equations relating
+collective operators to multi-local operators.
+
+On a final note, we consider the special case that all operators
+$O\in S$ are identical, for which we provide the expansions
+\begin{align}
+  N!_2\, X\otimes X \EQFS \p{\col{X}}^2 - \p{\col{X^2}},
+  &&
+  N!_3\, X^{\otimes 3} \EQFS \p{\col{X}}^3
+  - 3 \col{X} \odot \col{X^2} + 2\p{\col{X^3}},
+\end{align}
+\begin{align}
+  N!_4\, X^{\otimes 4}
+  \EQFS \p{\col{X}}^4 - 4 \p{\col{X}}^2\odot\p{\col{X^2}}
+  - 2 \col{X} \p{\col{X}}^2 \col{X}
+  + 3 \p{\col{X^2}}^2 + 8 \col{X} \odot \p{\col{X^3}}
+  - 6 \p{\col{X^4}}.
+\end{align}
+If furthermore $X^2=1$, as with e.g.~the Pauli operators of SU(2),
+then
+\begin{align}
+  N!_2\, X\otimes X \EQFS \p{\col{X}}^2 - N,
+  &&
+  N!_3\, X^{\otimes 3} \EQFS \p{\col{X}}^3 - \p{3N-2} \col{X},
+\end{align}
+\begin{align}
+  N!_4\, X^{\otimes 4} \EQFS \p{\col{X}}^4
+  - 2\p{3N-4} \p{\col{X}}^2 + 3 N \p{N-2}.
+\end{align}
 

### 2019-12-05 collective_sun/sun_perturbation.tex
fixed an error in the two-body perturbation theory

@@
 \DeclareMathOperator{\diag}{diag}
+\let\flat\relax
+\DeclareMathOperator{\flat}{flat}
 \DeclareMathOperator{\counts}{counts}
@@
   &&
-  \V_2 \equiv \f12 \sum_{p,q} w_{pq} X_p Y_q
+  \V_2 \equiv \sum_{p<q} w_{pq} \times \f12 \p{X_p Y_q + X_q Y_p}
   = \f12 \v X \c\m w \c\v Y,
@@
 \end{align}
-where $X,Y$ are trace-zero operators on an $n$-dimensional Hilbert
-space; $X_p,Y_p$ denote the action of $X,Y$ on spin $p$; $\v X,\v Y$
-are respectively vectors of all $X_p,Y_p$; $\v v$ and $\m w$ are
-respectively a vector and matrix of all $v_p,w_{pq}$; we enforce
-$w_{pq}=w_{qp}$ and $w_{pp}=0$; and we use a dot ($\c$) to denote an
-inner product between vectors and matrices as appropriate.
+where $X,Y$ are self-adjoint\footnote{In future versions of these
+  notes, we should relax the restriction that $X,Y$ are self-adjoint.}
+trace-zero operators on an $n$-dimensional Hilbert space; $X_p,Y_p$
+denote the action of $X,Y$ on spin $p$; $\v X,$ $\v Y$, $\v v$, and
+$\m w$ are respectively vectors and a matrix of all
+$X_p,Y_p,v_p,w_{pq}$; we use a dot ($\c$) to denote an inner product
+between vectors and matrices as appropriate; and we enforce
+$w_{pq}=w_{qp}$ and $w_{pp}=0$.
 
@@
   H_0 \V_1 \ket\psi
-  &= E_0 \V_1 \ket\psi
-  + \p{\sum_{p,q} h_{pq} v_p X_q - \sum_q v_q h_q X_q} \ket\psi \\
-  &= E_0 \V_1 \ket\psi
-  + \v v\c \p{\m h - \diag\v h}\c \v X \ket\psi,
+  = E_0 \V_1 \ket\psi
+  + \p{\sum_{p,q} h_{pq} v_p X_q - \sum_q h_q v_q X_q} \ket\psi,
   \label{eq:H_V1_psi}
 \end{align}
-where $\m h$ is a matrix of all coefficients $h_{pq}$ of the
-interaction Hamiltonian $H_0$ in \eqref{eq:ints}; $\v h$ is a vector
-of the net interaction energies $h_q\equiv\sum_ph_{pq}$ associated
-with individual spins $p$; and $\diag\v h$ is a matrix with $\v h$ on
-the diagonal and zeros everywhere else.  We thus find that the vector
-$\V_1\ket\psi$ is an eigenvector of the interaction Hamiltonian $H_0$
-if the coefficients $\v v$ satisfy
+where $h_q\equiv\sum_p h_{pq}$ is the net interaction energy
+associated with spin $p$.  We thus find that the vector $\V_1\ket\psi$
+is an eigenvector of the interaction Hamiltonian $H_0$ if the
+coefficients $\v v$ satisfy
 \begin{align}
-  \p{\m h - \diag\v h}\c\v v = \Delta\p{\v v} \v v.
+  \p{\m h - \diag\v h}\c\v v = \Delta\p{\v v} \v v,
   \label{eq:cond_1}
 \end{align}
-If we {\it construct} a perturbation $\V_1\p{\v v_\Delta}$ with
-coefficients $\v v_\Delta$ that satisfy \eqref{eq:cond_1} with
-eigenvalue $\Delta$, then
+where $\v h,\m h$ are respectively a vector and matrix of all
+$h_p,h_{pq}$; $\diag\v h$ is a matrix with $\v h$ on the diagonal and
+zeros everywhere else; and $\Delta\p{\v v}$ is some constant that may
+depend on $\v v$.  If we {\it construct} a perturbation
+$\V_1\p{\v v_\Delta}$ with coefficients $\v v_\Delta$ that satisfy
+\eqref{eq:cond_1} with eigenvalue $\Delta$, then
 \begin{align}
@@
   H_0 \V_2 \ket\psi
-  &= E_0 \V_2 \ket\psi
+  = E_0 \V_2 \ket\psi
   + \sum_{k,\ell} \bar h_{k\ell} w_{k\ell} X_k Y_\ell \ket\psi
-  + \sum_{k,\ell,q} h_{kq} w_{q\ell} X_k Y_\ell \ket\psi
-  - \sum_{k,\ell} w_{k\ell} h_{k\ell} X_k \odot Y_k \ket\psi \\
-  &= E_0 \V_2 \ket\psi
-  + \v X \c \p{\bar{\m h}\oc\m w + \m h\c\m w} \c \v Y \ket\psi
-  - \v\1 \c \p{\m h\oc\m w} \c \v{X\odot Y} \ket\psi,
+  + \f12 \sum_{k\ne\ell} \sum_q \p{h_{kq} w_{q\ell} + w_{kq} h_{q\ell}}
+  X_k Y_\ell \ket\psi,
   \label{eq:H_V2_psi}
@@
   &&
-  h_k \equiv \sum_\ell h_{k\ell};
+  h_k \equiv \sum_\ell h_{k\ell}.
 \end{align}
-$\m w,\m h,\bar{\m h}$ are matrices of all
-$w_{k\ell},h_{k\ell},\bar h_{k\ell}$; $\m x\oc\m y$ is an element-wise
-(Hadamard) product of $\m x$ and $\m y$, i.e.~with matrix elements
-$x_{k\ell} y_{k\ell}$; $\v\1\equiv\p{1,1,1,\cdots}$ is an
-$N$-component vector of ones; and $\v{X\odot Y}$ is a vector of all
-$X_k\odot Y_k$.  We thus find that the vector $\V_2\ket\psi$ is an
-eigenvector of the interaction Hamiltonian $H_0$ if the coefficient
-matrix $\m w$ of all $w_{pq}$ satisfies
+We thus find that the vector $\V_2\ket\psi$ is an eigenvector of the
+interaction Hamiltonian $H_0$ if the coefficients $\m w$ satisfy
 \begin{align}
-  \bar{\m h}\oc\m w + \m h\c\m w = \f12 \Delta\p{\m w} \m w,
+  2\bar{\m h} \circ \m w + \m h\c\m w + \m w\c\m h
+  = \Delta\p{\m w} \m w,
   \label{eq:cond_2}
 \end{align}
-for some constant $\Delta\p{\m w}$ that may depend on $\m w$.
-Remembering that the matrices $\m h$ and $\m w$ are both symmetric
-(under transposition) with zeros on the diagonal, the diagonal
-components of the condition in \eqref{eq:cond_2} automatically enforce
-that the last term ($\sim\v{X\odot Y}$) in \eqref{eq:H_V2_psi}
-vanishes.
+where $\m w,\m h,\bar{\m h}$ are matrices of all
+$w_{k\ell},h_{k\ell},\bar h_{k\ell}$; $\bar{\m h} \circ \m w$ is an
+element-wise (Hadamard) product of $\bar{\m h}$ and $\m w$, i.e.~with
+matrix elements $\bar h_{k\ell} w_{k\ell}$; and $\Delta\p{\m w}$ is
+some constant that may depend on $\m w$.
 
@@
   \m w = \sum_{p,q} w_{pq} \op{p}{q}
-  \to \v{\m w} \equiv \sum_{p,q} w_{pq} \ket{pq},
+  \to \v{\m w} \equiv \flat\m w \equiv \sum_{p,q} w_{pq} \ket{pq},
 \end{align}
-and defining the matrix
+where $\flat\m m$ is a flattening of the matrix $\m m$ into a vector
+$\vec{\m m}$, and defining the matrix
 \begin{align}
-  \tilde{\m h}
-  \equiv \sum_{p,q} \bar h_{pq} \op{pq}
-  + \sum_{p,q,\ell} h_{pq} \op{p\ell}{q\ell},
+  \check{\m h}
+  \equiv 2\diag\p{\flat\bar{\m h}}
+  + \m h\otimes\1_N + \1_N\otimes\m h
+  = 2 \sum_{p,q} \bar h_{pq} \op{pq}
+  + \sum_{p,q,\ell} h_{pq} \p{\op{p\ell}{q\ell} + \op{\ell p}{\ell q}},
   \label{eq:h_super_mat}
 \end{align}
-in terms of which the condition in \eqref{eq:cond_2} becomes the
-eigenvalue equation
+where $\1_D$ is a $D\times D$ identity matrix.  The condition in
+\eqref{eq:cond_2} is then equivalently the eigenvalue equation
 \begin{align}
-  \tilde{\m h} \c \v{\m w} = \f12 \Delta\p{\m w} \v{\m w}.
+  \p{\check{\m d} \c \check{\m h} \c \check{\m d}} \c \v{\m w}
+  = \Delta\p{\m w} \v{\m w},
+  &&
+  \check{\m d} \equiv \1_{N^2} - \sum_\ell \op{\ell\ell},
   \label{eq:cond_2_eig}
 \end{align}
-The matrix $\tilde{\m h}$ in \eqref{eq:h_super_mat} can be written in
-the block-diagonal form
-\begin{align}
-  \tilde{\m h} = \sum_\ell \m h_\ell \otimes \op{\ell},
-  &&
-  \m h_\ell
-  \equiv \sum_{p,q}\p{\delta_{pq} \bar h_{p\ell} + h_{pq}} \op{p}{q},
-\end{align}
-which reduces the eigenvalue equation in \eqref{eq:cond_2_eig} to the
-$N$ smaller, independent eigenvalue equations
-\begin{align}
-  \m h_\ell \c \v w_\ell = \f12 \Delta\p{\m w} \v w_\ell,
-  &&
-  \v w_\ell \equiv \sum_p w_{p\ell} \ket{p},
-  \label{eq:cond_2_eig_block}
-\end{align}
-where $\v w_\ell$ is essentially the $\ell$-th row or column of
-$\m w$.  We can (numerically) diagonalize each of the blocks
-$\m h_\ell$, and classify their eigenvectors $\v w_\ell$ by the
-corresponding eigenvalues $\Delta\p{\m w}/2$.  ``Total'' eigenvectors
-$\v{\m w}$ of $\tilde{\m h}$ are then built from eigenvectors
-$\v w_\ell$ within each block $\m h_\ell$ that share the same
-eigenvalue $\Delta\p{\m w}/2$.  The task of constructing eigenvectors
-$\V_2\p{\m w_\Delta}\ket\psi$ of $H_0$ and determining their energy
-thus reduces to finding the eigenvectors and eigenvalues of $N$
-matrices $\m h_\ell$ of size $N\times N$.
+where $\check{\m d}$ annihilates the diagonal part of a (flattened)
+matrix it acts on, while keeping the off-diagonal components the same.
+We conjugate $\check{\m h}$ by $\check{\m d}$ to enforce that its
+eigenvectors $\v{\m w}$ correspond to matrices $\m w$ with zeros on
+the diagonal, i.e.~$w_{pp}=0$.
 
-If the interaction Hamiltonian $H_0$ is translationally invariant,
-then $\bar h_{k\ell}=h_{k\ell}-h$, and $h_{k\ell}$ depends only on the
-displacement between the spins indexed by $k$ and $\ell$.  The blocks
-$\m h_\ell$ then take the form
-\begin{align}
-  \m h_{\ell} = \m\Sigma_\ell \c \m h_0 \c \m\Sigma_\ell^\dag,
-  &&
-  \m\Sigma_\ell \equiv \sum_p \op{p+\ell}{p},
-  \label{eq:block_shift}
-\end{align}
-which implies that we need only find the eigenvectors and eigenvalues
-of a single block $\m h_0$, after which we can determine the
-eigenvectors of any block simply by appropriate applications of the
-shift matrix $\m\Sigma_\ell$.  Note that in dimensions $D>1$, the
-relation in \eqref{eq:block_shift} holds for vectors $p,\ell$ that
-index lattice sites, i.e.~$p=\p{p_1,p_2,\cdots,p_D}$ and likewise with
-$\ell$.
+Both numerical experiments and physical arguments suggest that the
+$N^2\times N^2$ matrix $\check{\m d} \c \check{\m h} \c \check{\m d}$
+is highly degenerate, which means that we might be able to simplify
+\eqref{eq:cond_2_eig} even further.  We defer this simplification to
+future work, and for now bear the (large) computational cost of
+diagonalizing a $N^2\times N^2$ matrix.
 
@@
 coefficients $\m w_\Delta$ that satisfy \eqref{eq:cond_2_eig} with
-eigenvalue $\Delta/2$, then
+eigenvalue $\Delta$, then
 \begin{align}
@@
 matrix $\m w$ into its projections $\m w_\Delta$ onto the eigenspace
-of matrices satisfying \eqref{eq:cond_2_eig} with eigenvalue
-$\Delta/2$, i.e.~$\m w=\sum_\Delta\m w_\Delta$, which allows us to
-expand
+of matrices satisfying \eqref{eq:cond_2_eig} with eigenvalue $\Delta$,
+i.e.~$\m w=\sum_\Delta\m w_\Delta$, which allows us to expand
 \begin{align}
@@
 $\m w=\sum_\Delta\m w_\Delta$, where $\m w_\Delta$ is the projection
-of $\m w$ onto the $\p{\Delta/2}$-eigenspace of the operator
-$\tilde{\m h}$ defined in \eqref{eq:h_super_mat}.
+of $\m w$ onto the $\Delta$-eigenspace of the operator
+$\check{\m d}\c\check{\m h}\c\check{\m d}$ defined by
+\eqref{eq:h_super_mat} and \eqref{eq:cond_2_eig}.
 
@@
 which has one term with $\set{p,q}=\set{k,\ell}$, terms with
-$p,q\notin\set{k,\ell}$, and terms with $p,q\in\set{k,\ell}$.  The
-permutation operator $\Pi_{pq}$ acts trivially on
-$X_k Y_\ell \ket\psi$ when $\set{p,q}=\set{k,\ell}$ or
+$p,q\notin\set{k,\ell}$, and terms with only one of
+$p,q\in\set{k,\ell}$.  The permutation operator $\Pi_{pq}$ acts
+trivially on $X_k Y_\ell \ket\psi$ when $\set{p,q}=\set{k,\ell}$ or
 $p,q\notin\set{k,\ell}$, leaving a sum over $p,q$ of the form
@@
 \end{align}
-The terms in \eqref{eq:H_V2_psi_start} with only one of
-$p,q\in\set{k,\ell}$ simplify to
+Using similar techniques, we can simplify the terms in
+\eqref{eq:H_V2_psi_start} with only one of $p,q\in\set{k,\ell}$ to
 \begin{align}
-  \f12 \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}} w_{k\ell}
-  \p{h_{kq} X_q Y_\ell + h_{\ell q} X_k Y_q} \ket\psi
-  = \sum_{k,\ell,q} w_{k\ell} h_{kq} X_q Y_\ell \ket\psi
-  - \sum_{k,\ell} w_{k\ell} h_{k\ell} X_k\odot Y_k \ket\psi
+  \f12 \sum_{k\ne\ell} \sum_q \p{h_{kq} w_{q\ell} + w_{kq} h_{q\ell}}
+  X_k Y_\ell \ket\psi.
 \end{align}
-where $Z_k\equiv\p{X_kY_k+Y_kX_k}/2$.  In total,
+In total,
 \begin{align}
@@
   + \sum_{k,\ell} \bar h_{k\ell} w_{k\ell} X_k Y_\ell \ket\psi
-  + \sum_{k,\ell,q} h_{kq} w_{q\ell} X_k Y_\ell \ket\psi
-  - \sum_{k,\ell} w_{k\ell} h_{k\ell} X_k\odot Y_k \ket\psi.
+  + \f12 \sum_{k\ne\ell} \sum_q \p{h_{kq} w_{q\ell} + w_{kq} h_{q\ell}}
+  X_k Y_\ell \ket\psi.
 \end{align}
@@
 $O\in S$.  To this end, for any partition\footnote{A partition
-  $P\equiv\set{p}$ of $S$ consists of mutually disjoint, non-empty
+  $P\equiv\set{p}$ of $S$ is a set of mutually disjoint, non-empty
   subsets $p\subset S$ with $\bigcup_{p\in P}p=S$.}  $P\equiv\set{p}$
@@
 \end{align}
-The set of multi-local operators $\set{O_P:\t{partitions}~P~\t{of}~S}$
-consists of all operators that
+In words, a partition $P$ of $S$ consists of an assignment of
+operators $O\in S$ to individual parts $p\subset S$; each part $p$ is
+associated with a tensor factor of $O_P$, and the operators $O\in p$
+act on that tensor factor.  The set of multi-local operators
+$\set{O_P:\t{partitions}~P~\t{of}~S}$ consists of all operators that
 \begin{enumerate*}
@@
 \end{enumerate*}
-For any set $S$ of single-spin operators or partition $P$ of $S$, we
-can similarly define the collective operators
+For any set $S$ of single-spin operators and any partition $P$ of $S$,
+we can similarly define the collective operators
 \begin{align}
@@
 $\L$, and in turn solve $\L$ for $O_S$.  When
-$\abs{S}\in\set{1,2,3,4}$ (and, by conjecture, all $S$), by use of a
-computer algebra system (namely, Mathematica) we find that
+$\abs{S}\in\set{1,2,3,4}$ (and, by conjecture, for all $S$), by use of
+a computer algebra system (namely, Mathematica) we find that
 \begin{align}
@@
   \p{-1}^{\abs{S}-\abs{\tilde P}} \O_{\tilde P}
-  \times \prod_{p\in\tilde P} \p{\abs{p}-1}!
+  \times \prod_{\tilde p\in\tilde P} \p{\abs{\tilde p}-1}!
   \label{eq:multi_to_collective_full_simp}
@@
 (tuple) $\p{O_1,O_2,\cdots,O_{\abs{S}}}$ of all single-spin operators
-in $S$, and we enforce that partitions $\tilde P$ of an ordered list
-$\Pi$, as well as parts $p\in\tilde P$, are lexicographically ordered
-in accordance with the ordering of operators in $\Pi$.  If
+in $S$; and we enforce that partitions $\tilde P$ of an ordered list
+$\Pi$, as well as parts $\tilde p\in\tilde P$, are lexicographically
+ordered in accordance with the ordering of operators in $\Pi$.  If
 $S=\set{X,Y,Z}$, for example, then the ordered partitions of the
@@
   &&
-  O_{\tilde P} \equiv \bigotimes_{p\in\tilde P} \prod_{O\in p} O,
+  O_{\tilde P}
+  \equiv \bigotimes_{\tilde p\in\tilde P} \prod_{O\in\tilde p} O,
   &&
@@
   \O_{\tilde P}
-  \equiv \prod_{p\in\tilde P} \col{\textstyle\prod_{O\in p} O},
+  \equiv \prod_{\tilde p\in\tilde P}
+  \col{\textstyle\prod_{O\in\tilde p} O},
 \end{align}
-in which case we can also expand
+where products over elements of $\tilde S,\tilde P,\tilde p$ are taken
+in order according to the elements of these ordered lists, e.g.~if
+$\tilde S=\p{O_1,O_2,\cdots,O_{\abs{S}}}$, then
+$\prod_{O\in\tilde S}\equiv O_1O_2\cdots O_{\abs{S}}$.  These
+definitions allow us to expand
 \begin{align}

### 2019-12-06 collective_sun/sun_perturbation.tex
added appendix on a minor improvement to the second order two-body theory

@@
 \newcommand{\EQFS}{=_{\text{FS}}}
-\newcommand{\col}{\underline}
+\newcommand{\col}{\overline}
 
@@
 \DeclareMathOperator{\cov}{cov}
-\let\var\relax
+\let\var\oldvar
 \DeclareMathOperator{\var}{var}
 \DeclareMathOperator{\diag}{diag}
-\let\flat\relax
+\let\flat\oldflat
 \DeclareMathOperator{\flat}{flat}
 \DeclareMathOperator{\counts}{counts}
+\newcommand{\flatvec}{\underline\flat\,}
+\newcommand{\ul}{\underline}
 
@@
 \section{Single-body perturbations}
+\label{sec:single_body}
 
@@
 zeros everywhere else; and $\Delta\p{\v v}$ is some constant that may
-depend on $\v v$.  If we {\it construct} a perturbation
-$\V_1\p{\v v_\Delta}$ with coefficients $\v v_\Delta$ that satisfy
-\eqref{eq:cond_1} with eigenvalue $\Delta$, then
+depend on $\v v$.  Solving the eigenvalue problem in \eqref{eq:cond_1}
+requires diagonalizing an $N\times N$ matrix.  When the interaction
+Hamiltonian $H_0$ is translationally invariant, this eigenvalue
+problem can be solved analytically, which we discuss further in
+Appendix \ref{sec:trans_inv}.
+
+If we {\it construct} a perturbation $\V_1\p{\v v_\Delta}$ with
+coefficients $\v v_\Delta$ that satisfy \eqref{eq:cond_1} with
+eigenvalue $\Delta$, then
 \begin{align}
@@
 \section{Two-body perturbations}
+\label{sec:two_body}
 
@@
 \end{align}
-where $\flat\m m$ is a flattening of the matrix $\m m$ into a vector
-$\vec{\m m}$, and defining the matrix
+where $\flat\m m$ is a flattening of the $N\times N$ matrix $\m m$
+into an $N^2$-component vector $\v{\m m}$, and defining the
+$N^2\times N^2$ matrix
 \begin{align}
@@
 eigenvectors $\v{\m w}$ correspond to matrices $\m w$ with zeros on
-the diagonal, i.e.~$w_{pp}=0$.
-
-Both numerical experiments and physical arguments suggest that the
-$N^2\times N^2$ matrix $\check{\m d} \c \check{\m h} \c \check{\m d}$
-is highly degenerate, which means that we might be able to simplify
-\eqref{eq:cond_2_eig} even further.  We defer this simplification to
-future work, and for now bear the (large) computational cost of
-diagonalizing a $N^2\times N^2$ matrix.
+the diagonal, i.e.~$w_{pp}=0$.  Solving the eigenvalue equation in
+\eqref{eq:cond_2_eig} nominally requires diagonalizing an
+$N^2\times N^2$ matrix; we provide a reduction of this problem to that
+of diagonalizing an ${N\choose2}\times{N\choose2}$
+($\approx N^2/2\times N^2/2$ when $N\gg1$) matrix in Appendix
+\ref{sec:reduction}.  We also discuss how to reduce this problem even
+further in the case of a translationally invariant interaction
+Hamiltonian $H_0$ in Appendix \ref{sec:trans_inv}.
 
@@
 
+\newpage
 \appendix
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Reducing the two-body eigenvalue problem}
+\label{sec:reduction}
+
+At second order, the two-body perturbation theory developed in Section
+\ref{sec:two_body} nominally requires diagonalizing an $N^2\times N^2$
+matrix.  Here we discuss the reduction of this problem to that of
+diagonalizing an ${N \choose 2} \times {N \choose 2}$
+($\approx N^2/2\times N^2/2$ when $N\gg1$) matrix.  The basic idea is
+to remove redundancies and trivial values from the coefficient matrix
+$\m w$ that is used to construct a two-body perturbation
+$\V_2\p{\m w}$.  To this end, we flatten $\m w$ into a vector
+$\ul{\v{\m w}}$ in $\RR^{N\choose2}$,
+\begin{align}
+  \m w = \sum_{p,q} w_{pq} \op{p}{q}
+  \to \ul{\v{\m w}} \equiv \flatvec\m w
+  \equiv \sum_{p<q} w_{pq} \ket{\set{p,q}},
+\end{align}
+where we label basis vectors $\ket{\set{p,q}}$ by a choice of two
+distinct integers from $\set{1,2,\cdots,N}$.  We then define the
+matrix
+\begin{align}
+  \ul{\check{\m h}} = 2\diag\p{\flatvec\bar{\m h}}
+  + \sum_{p<q}
+  \sp{\sum_k \p{h_{pk} \ket{\set{q,k}} + h_{qk} \ket{\set{p,k}}}}
+  \bra{\set{p,q}}.
+\end{align}
+The condition in \eqref{eq:cond_2} is then equivalently the eigenvalue
+problem
+\begin{align}
+  \ul{\check{\m h}} \c \ul{\v{\m w}}
+  = \Delta\p{\m w} \ul{\v{\m w}}.
+\end{align}
+Unlike the eigenvalue problem in \eqref{eq:cond_2_eig}, there is now
+no need to conjugate $\ul{\check{\m h}}$ by an operator that
+annihilates the diagonal components of $\m w$, because the
+${N\choose2}$-component vector $\ul{\v{\m w}}$ already contains only
+the non-trivial parts of $\m w$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Making use of translational invariance}
+\label{sec:trans_inv}
+
+[Discuss simplifications that occur to the single- and two-body theory
+if $H_0$ is translationally invariant.]
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Collective and multi-local operators in the fully symmetric

### 2019-12-06 soc_squeezing/collective_paper/review/response.tex
first (incomplete) draft of response to reviewers on collective spin paper

\documentclass[preprint,superscriptaddress]{revtex4-2}

% linking references
\usepackage{hyperref}
\hypersetup{
  breaklinks=true,
  colorlinks=true,
  linkcolor=blue,
  urlcolor=cyan,
}

%%% symbols, notations, etc.
\usepackage{physics,braket,bm,commath,amssymb}
\renewcommand{\t}{\text} % text in math mode
\newcommand{\f}[2]{\dfrac{#1}{#2}} % shorthand for fractions
\newcommand{\p}[1]{\left( #1 \right)} % parenthesis
\renewcommand{\sp}[1]{\left[ #1 \right]} % square parenthesis
\renewcommand{\v}{\bm} % bold vectors
\newcommand{\uv}[1]{\bm{\hat{#1}}} % unit vectors

\renewcommand{\abs}[1]{\lvert #1 \rvert}

\newcommand{\bk}{\braket} % shorthand for braket notation
\newcommand{\Bk}{\Braket}

\newcommand{\ul}{\underline} % shorthand for underline

\newcommand{\A}{\mathcal{A}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\J}{\mathcal{J}}
\renewcommand{\O}{\mathcal{O}}
\newcommand{\Q}{\mathcal{Q}}
\renewcommand{\S}{\mathcal{S}}
\newcommand{\U}{\mathcal{U}}

\newcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathbb{N}}

\newcommand{\z}{\text{z}}
\newcommand{\x}{\text{x}}
\newcommand{\y}{\text{y}}
\newcommand{\Z}{\text{Z}}
\newcommand{\X}{\text{X}}
\newcommand{\Y}{\text{Y}}
\newcommand{\bmu}{{\bar\mu}}
\newcommand{\bnu}{{\bar\nu}}

\usepackage{dsfont} % for identity operator
\newcommand{\1}{\mathds{1}}

\newcommand{\up}{\uparrow}
\newcommand{\dn}{\downarrow}

\renewcommand{\a}{\alpha} % free index
\renewcommand{\b}{\beta} % free index

%%% figures
\usepackage{graphicx} % for figures
\usepackage[caption=false]{subfig} % subfigures (via \subfloat[]{})
\newcommand{\sref}[1]{\protect\subref{#1}} % for referencing subfigures
\graphicspath{{./figures/}} % set path for all figures

% remove extra white space inside enumerate environments
\usepackage[inline]{enumitem}
\setenumerate{label=(\roman*)}
\setlist[enumerate]{leftmargin=*}

\setlength{\parindent}{0pt}
\setlength{\parskip}{1ex}

% for colored text
\usepackage[dvipsnames]{xcolor}
\newcommand{\blue}[1]{{\color{blue} #1}}
\newcommand{\red}[1]{{\color{red} #1}}
\newcommand{\green}[1]{{\color{ForestGreen} #1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

Dear Editors,

Please find attached our revised manuscript AJ11834, titled
``Short-time expansion of Heisenberg operators in open collective
quantum spin systems'', for consideration for publication in Physical
Review A.  Below, we address the questions and comments made by our
referees, and detail the corresponding changes that we have made to
our manuscript.  We hope that our responses and changes address the
referees' questions to a satisfactory degree, leaving no more
outstanding questions or concerns about our work.

Excerpts from the referee reports are written in \blue{blue}, excerpts
from the previous version of our manuscript are written in \red{red},
and excerpts from the revised version of our manuscript are written in
\green{green}.  Unless explicitly stated otherwise, all equation
numbers, page numbers, etc.~refer to those in the revised manuscript.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{First referee}

\begin{enumerate}
\item Concerning the first referee's summary and comments:

  \blue{The authors have studied open collective spin systems with
    generic Markovian decoherence and presented a new method to
    calculate short-time correlators. The method is based on an
    expansion of Heisenberg operators and truncation of it for short
    times (TST). The number of particles that the presented method can
    deal with are shown to be rather high. The method is meant to work
    well with highly-correlated systems. Spin squeezing, two-time
    correlation functions, and out-of-time-ordered correlators are
    calculated as an application of the new method. As technical parts
    of the paper, authors calculate all the ingredients, e.g., the
    structure constants of a collective spin operator algebra, which
    are necessary to apply the method.}

  \blue{Although the developed method works only at short times, it
    can be fruitful for applications in the field of quantum
    metrology. One of the key points that makes the paper interesting
    is the capability of the TST expansion to compute the spin
    correlators in a regime (strong correlations) that other methods
    cannot practically access. In the regime that it works, the method
    is efficient and very promising for future applications.}

  \blue{Overall, the paper is well written and the presented method
    and following results deserve publication in Phys.~Rev.~A. In
    particular, I expect it to be utilized in the spin squeezed
    systems. I just have the following question:}

  We thank the referee for their summary of our work, and for the
  conclusion that our results deserve publication in Phys.~Rev.~A.  We
  similarly hope that our method finds useful applications in spin
  squeezed systems, as well as others.  We address the referee's
  question below.

\item Concerning the first referee's question:

  \blue{In the last paragraph of page 4, which continues to page 5,
    authors explain how unphysical result can be marked by and
    unphysical squeezing parameter. In the same paragraph authors list
    three possible mechanisms for breakdown. Is there a fundamental
    relationship between all of the three breakdown mechanisms and the
    value of the squeezing parameter? Asked differently, could it be
    possible that one of the three breakdown mechanisms happens
    without the squeezing parameter going unphysical?}

  The referee brings up a good question about diagnosing the breakdown
  of our method (TST).  In fact, there is {\it no} fundamental
  relationship between the breakdown of TST and an unphysical
  squeezing parameter ($\xi^2<0$), and it {\it is} possible for
  breakdown to occur without the squeezing parameter going unphysical.
  Instead, there is a direct relationship between the breakdown of TST
  and the value of individual correlators $\bk{\hat S_{\v n}}$ that
  are used to compute the squeezing parameter $\xi^2$.  The breakdown
  of TST affects (non-negligible) contributions to correlators
  $\bk{\hat\S_{\v n}}$ that grow with the time $t$ raised to some
  large power, so after breakdown these correlators rapidly take on
  unphysical values.  It is often the case that unphysical correlators
  will lead to an unphysical squeezing parameter, but not always.  The
  breakdown of TST should therefore be diagnosed by inspecting the
  correlators used to compute the squeezing parameter, rather than by
  inspecting the squeezing parameter itself.

  In order to clarify the diagnosis of breakdown for the TST
  expansion, we have rearranged and added text to the paragraph in
  question, replacing (end of page 4 through some of page 5 in the old
  text):

  \red{The main lesson from Figure 1 is that the TST expansion yields
    essentially exact results right up until a sudden and drastic
    departure that can be diagnosed by inspection.  {\bf The breakdown
      of the TST expansion in Figure 1 is marked by an unphysical
      squeezing parameter $\xi^2<0$, which occurs due to individual
      correlators taking unphysical values with
      $\abs{\bk{\hat\S_{\v n}\p{t}}}\gtrsim S^{\abs{\v n}}$.  Going up
      through order $M=70$ in the TST expansion does not significantly
      increase the breakdown time $t_{\t{break}}^{(M)}$, and in some
      cases even shortens $t_{\t{break}}^{(M)}$.}  The sudden and
    drastic departure from virtually exact results is consistent with
    the limitations of the TST expansion discussed at the end of
    Section II.  Specifically, we identify three possible mechanisms
    for breakdown:
    \begin{enumerate*}
    \item a rapid growth in the order $M$ necessary for the TST
      expansion to converge,
    \item the growth of numerical errors in excessively large terms of
      the TST expansion, and
    \item the formal breakdown of the perturbative expansion in the
      time $t$.
    \end{enumerate*}
    In all of these cases, a detailed cancellation eventually ceases
    to occur between large terms at high orders in the TST expansion.
    These large terms grow with the time $t$ raised to some large
    power (as high as $M$), and therefore rapidly yield wildly
    unphysical results.  In contrast to other approximate methods such
    as the cumulant expansion[58], the TST expansion can thus diagnose
    its own breakdown, which is an important feature when working in
    parameter regimes that are inaccessible via other means to compute
    correlators.}

  by (top of page 5 in the new text):

  \green{The main lesson from Figure 1 is that the TST expansion
    yields essentially exact results right up until a sudden and
    drastic departure that can be diagnosed by inspection.  {\bf The
      breakdown of the TST expansion in Figure 1 induces an unphysical
      squeezing parameter $\xi^2<0$.  In general, however, there is no
      fundamental relationship between the breakdown of the TST
      expansion and the conditions for a physical squeezing parameter
      $\xi^2$.  A proper diagnosis of breakdown therefore requires
      inspection of the correlators $\bk{\hat\S_{\v n}\p{t}}$ used to
      compute the squeezing parameter $\xi^2$, which upon breakdown
      will rapidly take unphysical values with
      $\abs{\bk{\hat\S_{\v n}\p{t}}}\gtrsim S^{\abs{\v n}}$.}  The
    sudden and drastic departure from virtually exact results is
    consistent with the limitations of the TST expansion discussed at
    the end of Section II.  Specifically, we identify three possible
    mechanisms for breakdown:
    \begin{enumerate*}
    \item a rapid growth in the order $M$ necessary for the TST
      expansion to converge,
    \item the growth of numerical errors in excessively large terms of
      the TST expansion, and
    \item the formal breakdown of the perturbative expansion in the
      time $t$.
    \end{enumerate*}
    In all of these cases, a detailed cancellation eventually ceases
    to occur between large terms at high orders in the TST expansion.
    These large terms grow with the time $t$ raised to some large
    power (as high as $M$), and therefore rapidly yield wildly
    unphysical results.  In contrast to other approximate methods such
    as the cumulant expansion[58], the TST expansion can thus diagnose
    its own breakdown, which is an important feature when working in
    parameter regimes that are inaccessible via other means to compute
    correlators.  {\bf Note that, due to the breakdown mechanisms of
      the TST expansion, going up through order $M=70$ does not
      significantly increase the breakdown time $t_{\t{break}}^{(M)}$
      in Figure 1, and in some cases even shortens
      $t_{\t{break}}^{(M)}$.}}

  Note that we used of bold text to help identify the changes above;
  the manuscript itself does not make these changes bold.  We hope
  that these changes clarify the relationship between breakdown of the
  TST expansion and an unphysical squeezing parameter.

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Second referee}

\begin{enumerate}
\item Concerning the second referee's summary and comments:

  \blue{In this work, the authors compute the dynamics of collective
    spin systems with decoherence via uncorrelated decay of individual
    spins by solving the operators’ Heisenberg’s equation of motion
    approximately.  It is carried out in the following way: Heisenberg
    operators are formally expanded into a Taylor series, which is
    then truncated to a finite order for short evolution times. The
    authors benchmark the method by computing spin squeezing, two-time
    correlation functions, and out-of-time-ordered correlators.

    The paper is clearly written, and the derivations appear to be
    sound.  However, there are some points need to be clarified before
    this paper can be accepted for publication.}

  We thank the referee for their succinct summary of our method, and
  for the the recognition that our manuscript is well written, and our
  derivations sound.  We address the refere's points below, and hope
  that the corresponding changes to our manuscript deem our work
  acceptable for publication in Physical Review A.

\item Concerning the first point, about novelty:

  \blue{1. The authors declare that they present a ‘new’
    method. However, it seems that the method as described resembles
    (if not the same as) the well-known Mori formalism, or memory
    function method, developed by Mori, Berne, Lee, Grigolini and
    others [see for example H.~Mori, Prog.~Theor.~Phys.~34 399
    (1965); James F Annett, et al.~Condens.~Matter 6, 645(1994)],
    where the Liouvillian superoperator is used to provide a compact
    representation of the time evolution of operators under
    Heisenberg’s equation of motion, and followed up with a continued
    fraction expansion. I hope the authors could acknowledge these
    previous studies and be careful with the claiming of ‘new’
    method.}

  We thank the referee for drawing our attention to the Mori formalism
  (\url{https://doi.org/10/cgtkbb}) and the related work by Annett et
  al.~(\url{https://doi.org/10/d79ngr}).  Indeed, there are some
  interesting connections between these works and ours, as pointed out
  by the referee.  For example, all of these methods rely on some sort
  of formal solution to the equations of motion for Heisenberg
  operators.  Nonetheless, we strongly believe that our methods are
  substantively distinct from those of these past works, and expand on
  some differences below.  In acknowledgment of the resemblance and
  connections to past works, however, we have also followed the
  referee's suggestion and changed all instances of ``\red{a new
    method}'' in our manuscript to ``\green{an efficient method}''.
  We have also added references to the Mori formalism and follow-up
  works (outlined below).

  At a high level, the truncated short-time (TST) expansion for a
  Heisenberg operator, Eq.~(10) in our manuscript, essentially
  involves only products of (i) initial-time expectation values of
  operators chosen from a fixed basis, and (ii) matrix elements of the
  time derivative operator $T$ (essentially the Liouvillian), as well
  as powers of $T$.  In order to evaluate the TST expansion, our
  method makes use of the structure constants of a collective spin
  operator algebra.  The Mori formalism, meanwhile, arrives at an
  expansion of Heisenberg operators (see Eq.~2.29 of
  \url{https://doi.org/10/cgtkbb}) that involves (i) Laplace
  transforms, (ii) operators, $f_j$, that are constructed recursively
  from the initial Heisenberg operator of interest, and (iii)
  coefficients, $C_j$, that are defined using the operators $f_j$ and
  additional Laplace transforms.  In order to evaluate this expansion,
  the Mori formalism provides a continued-fraction representation of
  the coefficients $C_j$.  Perhaps additional work might illuminate
  the connections between these two methods, but at face value they
  are quite different.

  Some of the work by Annett et al.~in \url{https://doi.org/10/d79ngr}
  has a closer resemblance to ours, in that Eq.~(25) therein is,
  similar to the TST expansion, a series expansion organized in powers
  of the time $t$.  Even here, however, Eq.~(25) concerns only one
  component of an (initially localized) Heisenberg operator $Q(t)$,
  namely the projection of $Q(t)$ onto $Q(0)$.  Furthermore, the
  solution of Eq.~(25) involves resolving $Q(t)$ into frequency
  components $R(\omega)$, representing the resolvents $R(\omega)$ by a
  continued fraction with some free parameters $\set{a_n,b_n}$, and
  then solving for the parameters $\set{a_n,b_n}$.  Our TST expansion
  considers explicitly non-local operators, solves for the entirety of
  these operators (as opposed to only one component), and involves
  neither frequency components nor continued fractions.  We hope that
  these considerations clarify how the TST expansion is distinct from
  past works.

  Despite these differences, we acknowledge that there may be some
  interesting connections between the above works and ours, and that
  investigating these connections may lead to a refinement of our
  work.  We have therefore changed the following sentence in the
  introduction (top left of page 2):

  \red{Our method is based on a short-time expansion of exact
    solutions to the equations of motion for Heisenberg operators.
    Evaluating this expansion requires ...}

  to:

  \green{Our method is based on a formal solution to the equations of
    motion for Heisenberg operators, thereby bearing some resemblance
    to the Mori formalism[59] and related work[60].  Specifically, we
    expand a formal solution for a Heisenberg operator into a Taylor
    series whose truncation can yield negligible error at sufficiently
    short times.  Evaluating the resulting expansion requires ...}

  We have also added a reference to these works to the discussion of
  the limitations of the TST expansion at the end of the theory
  section of our manuscript (right column of page 3), changing the
  sentence:

  \red{Precisely characterizing the implications of these last two
    considerations for the TST expansion requires additional analysis
    that we defer to future work.}

  to:

  \green{Precisely characterizing the implications of these last two
    considerations for the TST expansion requires additional analysis
    that we defer to future work, which may benefit from an
    investigation of the connections between the TST expansion and
    past works related to the Mori formalism[59, 60].}

  We hope that the referee considers these additions a sufficient
  acknowledgment of the Mori formalism and related work.

\item Concerning the second question, about decoherence:

  \blue{2. I am not sure if the paper properly takes into account
    previous results about the influence of decoherence on spin
    squeezing. First, there exists many mechanisms for decoherence,
    while decoherence considered in this paper is due to uncorrelated
    decay of individual spins. This is not clarified in the
    introduction. In fact, atom loss represents an important mechanism
    of decoherence in spin squeezing experiments, but is not
    considered in this paper. Could the method studied here also be
    applied to atom loss? It would be helpful if this point is
    mentioned. Second, for the specific decoherence model considered,
    the authors only cite a few studies about the influence of such
    decoherence on spin squeezing from one-axis twisting (OAT). There
    are many studies about the effect of decoherence on spin
    squeezing.}

  The referee makes a fair point that we claim to present a method for
  simulating ``generic'' Markovian decoherence, when in fact we only
  address certain (typical) forms of Markovian decoherence.  To avoid
  misleading readers about the contents of our paper, we have
  therefore replaced the words ``\red{generic Markovian decoherence}''
  by ``\green{typical forms of Markovian decoherence}'' in the
  abstract, as well as the last paragraph of the introduction.

  We would like to point out that our work {\it does} address more
  than just uncorrelated decay of individual spins: we also address
  uncorrelated excitation and dephasing of individual spins, as well
  as arbitrary coherent or incoherent combinations of these processes.
  Similarly, we provide all ingredients necessary to account for
  arbitrary combinations of {\it collective} decay, excitation, and
  dephasing in Appendices D-G, as mentioned in the paragraph below
  Eq.~(4) of our manuscript.

  The referee is correct, though, that our original manuscript does
  not discuss atom loss, which is an important decoherence mechanism
  in some experimental realizations of collective spin models.  We
  have therefore added the following discussion about atom loss to the
  theory section our manuscript (bottom right of page 2):

  \green{We note that particle loss is an important decoherence
    mechanism in many experimental realizations of collective spin
    models[41].  In principle, a spin model has no notion of the
    particle annihilation operators that generate particle loss, and
    therefore cannot capture this effect directly.  Nonetheless, for a
    system initially composed of $N$ particles, the effect of particle
    loss can be emulated with $O(1/N)$ error by the dissipator
    $\check\D_{\t{loss}}$ defined by
    $\check\D_{\t{loss}}\hat\S_{\v m}=-\abs{\v m}\hat\S_{\v m}$, where
    $\abs{\v m}\equiv\sum_\alpha m_\alpha$ (see Appendix H).
    Furthermore, the effect particle loss can be accounted for exactly
    by (i) introducing an additional index on spin operators,
    $\hat\S_{\v m}\to\hat\S_{N\v m}$, to keep track of different
    sectors of fixed particle number within a multi-particle Fock
    space, and (ii) constructing jump operators that appropriately
    couple spin operators within different particle-number sectors.
    We defer a detailed exact accounting of particle loss to future
    work.}

  The new appendix provides mathematical justification for our claim
  that the dissipator $\check\D_{\t{loss}}$ emulates particle loss
  with $O(1/N)$ error.  We hope that these additions do a satisfactory
  job of addressing particle loss.

  As a last point, concerning the number of references in our work to
  past studies about the influence of decoherence on spin squeezing in
  the one-axis twisting (OAT) model: the small number of such
  references is due to the fact that we do not have any particular
  discussions of the influence of decoherence on spin squeezing with
  OAT, and do not consider such discussions relevant to our current
  work.  Even with decoherence (of the types considered in our work),
  the OAT model is an exactly analytically solvable (see
  \url{https://doi.org/gcsjdp}).  Approximate numerical methods such
  as that presented in our manuscript are therefore inappropriate
  tools for the study of OAT.  The primary role of OAT in our work is
  to serve as a simple and exactly solvable benchmark of our numerical
  methods, rather than an object of study in and of itself.  To
  clarify this point, we have changed the second paragraph of section
  III on spin squeezing, benchmarking, and breakdown (top of page 4),
  from:

  \red{Note that the OAT model is a special case of the zero-field
    Ising model, whose quantum dynamics admits an exact analytic
    solution even in the presence of decoherence[62].  We will
    benchmark our calculations using these analytical OAT results
    wherever applicable (see Appendix J, as well as the Supplementary
    Material of Ref.~[14]).}

  to:

  \green{Note that the OAT model is a special case of the zero-field
    Ising model, whose quantum dynamics admits an exact analytic
    solution even in the presence of decoherence[64].  The approximate
    and numerics-oriented TST expansion is therefore ill-suited for
    studying the OAT model, which will merely serve as an exactly
    solvable benchmark of our methods.  Wherever applicable, we will
    provide exact results for the OAT model (see Appendix K, as well
    as the Supplementary Material of Ref.~[14]).}

  We hope that the referee and editors understand our position on this
  matter -- namely, that we would rather avoid adding spurious
  references to past works on a subject (the influence of decoherence
  on spin squeezing in the exactly solvable OAT model) that is of
  little relevance to our work on a numerical method to compute spin
  correlators.

\item Concerning the final question, about the robustness of the TNT
  model:

  \blue{3. I am surprised by the results from the strong decoherence
    regimes in Fig.~2, which shows that twist and-turn (TNT) model is
    much more robust than OAT and TAT. Is this physical or just an
    artifact from numerical errors. To rule out the latter, I suggest
    the authors check the results in the strong decoherence regime by
    studying a smaller system, where Monte Carlo simulation is
    accessible and can be used as a comparison. If this result is
    indeed physical, if would be very helpful if the authors could
    comment on the reason?}

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Miscellaneous changes}

\begin{enumerate}
\item In our discussions about the breakdown of the TST expansion in
  Sections II and III, we use the word ``unphysical'' to essentially
  mean ``not physically possible''.  In Section I, meanwhile, we
  comment about a method of simulation called the cumulant expansion,
  which breaks down in the presence of genuinely multi-body
  correlations.  In such a case, the cumulant expansion yields results
  that are {\it physically possible}, but {\it incorrect}.  For
  consistent use of language, we have therefore changed the following
  sentence in section I (bottom right of page 1):

  \red{The growth of genuinely multi-body correlations, however,
    eventually causes the cumulant expansion to yield unphysical
    results with no clear signature of failure.}

  to:

  \green{The growth of genuinely multi-body correlations, however,
    eventually causes the cumulant expansion to yield incorrect
    results with no clear signature of failure.}

\item One of our previous changes to the manuscript defined
  ``\green{$\abs{\v m}\equiv\sum_\alpha m_\alpha$}'' (bottom of page 2
  in the new text).  We have therefore removed a subsequent definition
  of this notation, which previously appeared in the following
  paragraph, ``\red{where
    $\abs{\v\ell}\equiv\sum_\alpha\ell_\alpha$}'' (middle right column
  of page 2 in the old text).

\end{enumerate}

\end{document}

### 2019-12-11 collective_sun/sun_perturbation.tex
simplified two-body eigenvalue problem in the case of translational invariance

--- 2019-12-05 collective_sun/sun_perturbation.tex
@@
 \renewcommand{\SS}{\mathbb{S}}
+\newcommand{\VV}{\mathbb{V}}
 \newcommand{\ZZ}{\mathbb{Z}}
-
-\newcommand{\PP}{\mathbb{Z}}
 
@@
 \DeclareMathOperator{\cov}{cov}
-\let\var\relax
+\let\var\oldvar
 \DeclareMathOperator{\var}{var}
 \DeclareMathOperator{\diag}{diag}
-\let\flat\relax
+\let\flat\oldflat
 \DeclareMathOperator{\flat}{flat}
 \DeclareMathOperator{\counts}{counts}
+\newcommand{\flatvec}{\underline\flat\,}
+\newcommand{\ul}{\underline}
+
+\usepackage{accents}
+\newcommand{\ut}{\undertilde}
+
+\newcommand{\floor}[1]{\lfloor{#1}\rfloor}
+\newcommand{\ceil}[1]{\lceil{#1}\rceil}
 
@@
 zeros everywhere else; and $\Delta\p{\v v}$ is some constant that may
-depend on $\v v$.  If we {\it construct} a perturbation
-$\V_1\p{\v v_\Delta}$ with coefficients $\v v_\Delta$ that satisfy
-\eqref{eq:cond_1} with eigenvalue $\Delta$, then
+depend on $\v v$.  Solving the general eigenvalue problem in
+\eqref{eq:cond_1} requires diagonalizing an $N\times N$ matrix.  When
+the interaction Hamiltonian $H_0$ is translationally invariant,
+however, this eigenvalue problem can be solved analytically; we
+provide the corresponding solution in Appendix \ref{sec:trans_inv}.
+
+If we {\it construct} a perturbation $\V_1\p{\v v_\Delta}$ with
+coefficients $\v v_\Delta$ that satisfy \eqref{eq:cond_1} with
+eigenvalue $\Delta$, then
 \begin{align}
@@
   = E_0 \V_2 \ket\psi
-  + \sum_{k,\ell} \bar h_{k\ell} w_{k\ell} X_k Y_\ell \ket\psi
-  + \f12 \sum_{k\ne\ell} \sum_q \p{h_{kq} w_{q\ell} + w_{kq} h_{q\ell}}
-  X_k Y_\ell \ket\psi,
+  + \f12 \sum_{p\ne q} \sp{2\bar h_{pq} w_{pq}
+    + \sum_k \p{h_{pk} w_{kq} + w_{pk} h_{kq}}}
+  X_p Y_q \ket\psi,
   \label{eq:H_V2_psi}
@@
 \begin{align}
-  \bar h_{k\ell} \equiv h_{k\ell} - \f12\p{h_k + h_\ell},
-  &&
-  h_k \equiv \sum_\ell h_{k\ell}.
+  \bar h_{pq} \equiv h_{pq} - \f12\p{h_p + h_q},
+  &&
+  h_p \equiv \sum_q h_{pq}.
 \end{align}
@@
 \begin{align}
-  2\bar{\m h} \circ \m w + \m h\c\m w + \m w\c\m h
-  = \Delta\p{\m w} \m w,
+  2 \bar h_{pq} w_{pq}
+  + \sum_{k\notin\set{p,q}} \p{h_{pk} w_{kq} + w_{pk} h_{kq}}
+  = \Delta\p{\m w} w_{pq},
   \label{eq:cond_2}
 \end{align}
-where $\m w,\m h,\bar{\m h}$ are matrices of all
-$w_{k\ell},h_{k\ell},\bar h_{k\ell}$; $\bar{\m h} \circ \m w$ is an
-element-wise (Hadamard) product of $\bar{\m h}$ and $\m w$, i.e.~with
-matrix elements $\bar h_{k\ell} w_{k\ell}$; and $\Delta\p{\m w}$ is
-some constant that may depend on $\m w$.
-
-We can recast the condition in \eqref{eq:cond_2} as an eigenvalue
-problem by writing the coefficient matrix $\m w$ as a vector
-$\v{\m w}$ in $\RR^N\times\RR^N$,
+where $\Delta\p{\m w}$ is some constant that may depend on $\m w$, and
+for clarity we explicitly exclude terms with $k\in\set{p,q}$ that
+vanish due to the fact that $\m h$ and $\m w$ are zero on the
+diagonal.  We can recast the condition in \eqref{eq:cond_2} as an
+eigenvalue problem by writing the coefficient matrix $\m w$ as a
+vector $\v{\m w}\in\RR^{N\choose 2}$,
 \begin{align}
   \m w = \sum_{p,q} w_{pq} \op{p}{q}
-  \to \v{\m w} \equiv \flat\m w \equiv \sum_{p,q} w_{pq} \ket{pq},
-\end{align}
-where $\flat\m m$ is a flattening of the matrix $\m m$ into a vector
-$\vec{\m m}$, and defining the matrix
-\begin{align}
-  \check{\m h}
-  \equiv 2\diag\p{\flat\bar{\m h}}
-  + \m h\otimes\1_N + \1_N\otimes\m h
-  = 2 \sum_{p,q} \bar h_{pq} \op{pq}
-  + \sum_{p,q,\ell} h_{pq} \p{\op{p\ell}{q\ell} + \op{\ell p}{\ell q}},
+  \to \v{\m w} \equiv \flat\m w
+  \equiv \sum_{p<q} w_{pq} \ket{\set{p,q}},
+\end{align}
+where we label basis of vectors $\ket{\set{p,q}}\in\RR^{N\choose 2}$
+by a choice of two distinct integers $p,q\in\ZZ_N$.  We similarly
+define the ${N \choose 2}\times{N \choose 2}$ matrix
+\begin{align}
+  \check{\m h} = 2\sum_{p<q} \bar h_{pq} \op{\set{p,q}}
+  + \sum_{p<q} \ket{\set{p,q}} \sum_{k\notin\set{p,q}}
+  \p{h_{pk} \bra{\set{k,q}} + h_{kq} \bra{\set{p,k}}}.
   \label{eq:h_super_mat}
 \end{align}
-where $\1_D$ is a $D\times D$ identity matrix.  The condition in
-\eqref{eq:cond_2} is then equivalently the eigenvalue equation
-\begin{align}
-  \p{\check{\m d} \c \check{\m h} \c \check{\m d}} \c \v{\m w}
-  = \Delta\p{\m w} \v{\m w},
-  &&
-  \check{\m d} \equiv \1_{N^2} - \sum_\ell \op{\ell\ell},
+The condition in \eqref{eq:cond_2} is then equivalently the eigenvalue
+equation
+\begin{align}
+  \check{\m h} \c \v{\m w} = \Delta\p{\m w} \v{\m w}.
   \label{eq:cond_2_eig}
 \end{align}
-where $\check{\m d}$ annihilates the diagonal part of a (flattened)
-matrix it acts on, while keeping the off-diagonal components the same.
-We conjugate $\check{\m h}$ by $\check{\m d}$ to enforce that its
-eigenvectors $\v{\m w}$ correspond to matrices $\m w$ with zeros on
-the diagonal, i.e.~$w_{pp}=0$.
-
-Both numerical experiments and physical arguments suggest that the
-$N^2\times N^2$ matrix $\check{\m d} \c \check{\m h} \c \check{\m d}$
-is highly degenerate, which means that we might be able to simplify
-\eqref{eq:cond_2_eig} even further.  We defer this simplification to
-future work, and for now bear the (large) computational cost of
-diagonalizing a $N^2\times N^2$ matrix.
+When the interaction Hamiltonian $H_0$ and perturbation $\V_2\p{\m w}$
+are translationally invariant, this eigenvalue problem can be reduced
+to that of diagonalizing a matrix of dimensions
+$\floor{N/2}\times\floor{N/2}$, as we show in Appendix
+\ref{sec:trans_inv}.
 
@@
 neglected scalar terms in $H_2^{(2)}$ without physical consequence.
-In order to expand the coefficients $W_k\p{\m h,\m w}$, for any matrix
-$\m m$ we define the vectors
-\begin{align}
-  \v{\m m} \equiv \sum_{p,q} m_{pq} \ket{pq},
+In order to expand the coefficients $W_k\p{\m h,\m w}$, for any
+symmetric matrix $\m m$ with zeroes on the diagonal we define the
+vectors
+\begin{align}
+  \v{\m m} \equiv \flat\m m \equiv \sum_{p<q} m_{pq} \ket{\set{p,q}},
   &&
@@
 \end{align}
-using which we have
+in terms of which
 \begin{align}
@@
   \equiv \sum_\Delta \f1\Delta \f{\v w_\Delta\c\v w_\Delta
-    - \v{\m w}_\Delta\c\v{\m w}_\Delta/2}{N!_4},
+    - \v{\m w}_\Delta\c\v{\m w}_\Delta}{N!_4},
 \end{align}
@@
   \equiv \sum_\Delta \f1{\Delta} \sp{\f{\v w_\Delta\c\v w_\Delta
-      - \v{\m w}_\Delta\c\v{\m w}_\Delta}{4N!_2}
+      - 2\v{\m w}_\Delta\c\v{\m w}_\Delta}{4N!_2}
     + \f{\p{3N-4}\p{\v w_\Delta\c\v w_\Delta
-        - \v{\m w}_\Delta\c\v{\m w}_\Delta/2}}{2N!_4}},
+        - \v{\m w}_\Delta\c\v{\m w}_\Delta}}{2N!_4}},
 \end{align}
@@
 $\m w=\sum_\Delta\m w_\Delta$, where $\m w_\Delta$ is the projection
-of $\m w$ onto the $\Delta$-eigenspace of the operator
-$\check{\m d}\c\check{\m h}\c\check{\m d}$ defined by
-\eqref{eq:h_super_mat} and \eqref{eq:cond_2_eig}.
-
+of $\m w$ onto the $\Delta$-eigenspace of the operator $\check{\m h}$
+defined in \eqref{eq:h_super_mat}.
+
+\newpage
 \appendix
@@
   H_0 \V_2 \ket\psi
-  = E_0 \V_2 \ket\psi
+  &= E_0 \V_2 \ket\psi
   + \sum_{k,\ell} \bar h_{k\ell} w_{k\ell} X_k Y_\ell \ket\psi
   + \f12 \sum_{k\ne\ell} \sum_q \p{h_{kq} w_{q\ell} + w_{kq} h_{q\ell}}
-  X_k Y_\ell \ket\psi.
+  X_k Y_\ell \ket\psi \\
+  &= E_0 \V_2 \ket\psi
+  + \f12 \sum_{k\ne\ell} \sp{2\bar h_{k\ell} w_{k\ell}
+    + \sum_q \p{h_{kq} w_{q\ell} + w_{kq} h_{q\ell}}}
+  X_k Y_\ell \ket\psi
 \end{align}
@@
   w_{k\ell} w_{pq} X_k Y_\ell X_p Y_q
-  \EQFS \v{\m w}\c\v{\m w} \p{X_0^2 Y_1^2 + X_0 Y_0 Y_1 X_1},
-  &&
-  \v{\m w} \equiv \sum_{k,\ell} w_{k\ell} \ket{k\ell}.
+  \EQFS 2 \v{\m w}\c\v{\m w} \p{X_0^2 Y_1^2 + X_0 Y_0 Y_1 X_1},
+  &&
+  \v{\m w} \equiv \sum_{k<\ell} w_{k\ell} \ket{\set{k,\ell}}.
 \end{align}
@@
   = \sum_{k,\ell,p} w_{k\ell} w_{kp} - \sum_{k,\ell} w_{k\ell}^2
-  = \v w\c\v w - \v{\m w}\c\v{\m w},
+  = \v w\c\v w - 2 \v{\m w}\c\v{\m w},
   &&
@@
   w_{k\ell} w_{pq}
-  = 2 \v{\m w}\c\v{\m w} - 4 \v w\c\v w.
+  = -4 \v w\c\v w + 4 \v{\m w}\c\v{\m w}.
 \end{align}
@@
   \P_0 \V_2^2 \P_0
-  \EQFS -\p{\v w\c\v w - \f12 \v{\m w}\c\v{\m w}} X_0 X_1 Y_2 Y_3
-  + \f14 \v{\m w}\c\v{\m w} \p{X_0^2 Y_1^2 + X_0 Y_0 Y_1 X_1} \\
-  + \f14 \p{\v w\c\v w - \v{\m w}\c\v{\m w}}
+  \EQFS -\p{\v w\c\v w - \v{\m w}\c\v{\m w}} X_0 X_1 Y_2 Y_3
+  + \f12 \v{\m w}\c\v{\m w} \p{X_0^2 Y_1^2 + X_0 Y_0 Y_1 X_1} \\
+  + \f14 \p{\v w\c\v w - 2\v{\m w}\c\v{\m w}}
   \p{X_0^2 Y_1 Y_2 + X_0 Y_0 X_1 Y_2
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Analytical results in translationally invariant systems}
+\label{sec:trans_inv}
+
+Here we simplify the eigenvalue problems that appear at second order
+in the single- and two-body theory in Sections \ref{sec:single_body}
+and \ref{sec:two_body} in the case of a translationally invariant
+system with periodic boundary conditions.  We solve the single-body
+eigenvalue problem analytically, and reduce the two-body eigenvalue
+problem to that of diagonalizing a single $\p{N-1}\times\p{N-1}$
+matrix.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Single-body eigenvalue problem}
+\label{sec:trans_inv_single}
+
+For reference, the single-body eigenvalue problem is
+\begin{align}
+  \p{\m h - \diag\v h}\c\v v = \Delta\p{\v v} \v v,
+  \label{eq:cond_1_ref}
+\end{align}
+where $\m h$ is a matrix of all couplings $h_{pq}$ that appear in the
+interaction Hamiltonian $H_0$ in \eqref{eq:ints}; $\v h$ is a vector
+of all $h_p\equiv\sum_q h_{pq}$; $\diag\v h$ is a matrix with $\v h$
+on the diagonal and zeroes everywhere else; and $\v v$ is a vector of
+coefficients used to construct the single-body perturbation
+$\V_1\p{\v v}$ defined in \eqref{eq:perturbations}.  When the
+interaction Hamiltonian $H_0$ is translationally invariant, the
+couplings $h_{pq}$ depend only on the separation $\abs{p-q}$, so the
+eigenvectors of $\m h$ are plane waves of the form
+\begin{align}
+  \v v_k \equiv \sum_p e^{ipk} \ket{p},
+\end{align}
+where $k\in\ZZ_N\times2\pi/N$.  The corresponding eigenvalues can be
+determined by expanding
+\begin{align}
+  \m h\c\v v_k = \sum_{p,q} h_{pq} e^{iqk} \ket{p}
+  = \sum_{p,d} h_{p,d+p} e^{i\p{d+p}k} \ket{p}
+  = \sum_d h_{0,d} e^{idk} \sum_p e^{ipk}\ket{p}
+  = \sum_d h_{0,d} \cos\p{dk} \v v_k.
+\end{align}
+The remainder of \eqref{eq:cond_1_ref} that we need to sort out is
+$\diag\v h$, where all $h_p=\sum_qh_{pq}=\sum_dh_{0,d}$ are equal,
+which implies that $\diag\v h=\sum_dh_{0,d}$ is a scalar.  We thus
+find that
+\begin{align}
+  \p{\m h - \diag\v h}\c\v v_k = \Delta_k \v v_k,
+  &&
+  \Delta_k \equiv \sum_d h_{0,d} \sp{\cos\p{dk}-1}.
+\end{align}
+Note that $\Delta_k=\Delta_{-k}$, which implies that there are at most
+$\lceil N/2\rceil$ distinct eigenvalues of $\m h - \diag\v h$ (one of
+which is $\Delta_0=0$).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Two-body eigenvalue problem}
+\label{sec:trans_inv_two}
+
+The two-body eigenvalue equation in \eqref{eq:cond_2_eig} is designed
+to find solutions $\m w$ to the equation
+\begin{align}
+  \p{2h_{pq}-h_p-h_q} w_{pq}
+  + \sum_{k\notin\set{p,q}} \p{h_{pk} w_{kq} + w_{pk} h_{kq}}
+  = \Delta w_{pq},
+  \label{eq:cond_2_ref_start}
+\end{align}
+where $h_p\equiv\sum_q h_{pq}$.  If the interaction Hamiltonian $H_0$
+and perturbation $\V_2\p{\m w}$ are translationally invariant, then
+the couplings $h_{pq}$ and coefficients $w_{pq}$ depend only on the
+separation $\abs{p-q}$.  In this case, all $h_p=h_0$ are equal, and
+\eqref{eq:cond_2_ref_start} can be written in the form
+\begin{align}
+  2 \m h\circ \m w + \m h\c\m w + \m w\c\m h
+  = \sp{\Delta + 2h_0} \m w,
+  \label{eq:cond_2_ref}
+\end{align}
+where $\m h\circ\m w$ denotes an element-wise (Kronecker) product of
+$\m h$ and $\m w$, i.e.~with matrix elements $h_{pq}w_{pq}$.  In order
+to find solutions to \eqref{eq:cond_2_ref}, we we expand the matrices
+$\m h,\m w$ in a basis that respects translational symmetry.  To this
+end, for all $d\in\ZZ_N$ we define the {\it shift matrices}
+\begin{align}
+  \m s_d \equiv \sum_p \op{p+d}{p},
+\end{align}
+where the sum $p+d$ is implicitly taken modulo $N$.  The shift
+matrices allow us to expand
+\begin{align}
+  \m h = \sum_{p,q} h_{pq} \op{p}{q}
+  = \sum_{d,q} h_{q+d,q} \op{q+d}{q}
+  = \sum_d h_{d,0} \m s_d,
+  &&
+  \m w = \sum_d w_{d,0} \m s_d,
+\end{align}
+and write the condition in \eqref{eq:cond_2_ref} as
+\begin{align}
+  \sum_{d\ne 0} h_{d,0} w_{d,0} \m s_d
+  + \sum_{\substack{c,d\ne0\\c+d\ne0}} h_{c,0} w_{d,0} \m s_{c+d}
+  = \p{\f12\Delta + h_0} \sum_{d\ne 0} w_{d,0} \m s_d,
+\end{align}
+where we explicitly exclude terms that always vanish due to the fact
+that $\m h,\m w$ are zero on the diagonal.  With some re-indexing, we
+can write
+\begin{align}
+  \sum_{d\ne 0} h_{d,0} w_{d,0} \m s_d
+  + \sum_{\substack{c,d\ne0}} h_{d-c,0} w_{c,0} \m s_d
+  = \p{\f12\Delta + h_0} \sum_{d\ne 0} w_{d,0} \m s_d,
+  \label{eq:cond_2_shift}
+\end{align}
+and define
+\begin{align}
+  \ul{\check{\m h}}
+  \equiv \sum_{d\ne0} h_{d,0} \op{d}
+  + \sum_{c,d\ne0} h_{d-c,0} \op{d}{c},
+  &&
+  \ul{\v{\m w}} \equiv \sum_{d\ne0} w_{d,0} \ket{d},
+\end{align}
+in order to express \eqref{eq:cond_2_shift} as the eigenvalue
+equation
+\begin{align}
+  \ul{\check{\m h}} \c \ul{\v{\m w}}
+  = \p{\f12\Delta + h_0} \ul{\v{\m w}}.
+  \label{eq:cond_2_square}
+\end{align}
+We can reduce this eigenvalue problem even further by making use of a
+reflection symmetry that comes from the fact that $h_{pq}=h_{qp}$ and
+$w_{pq}=w_{qp}$, which implies $h_{d,0}=h_{-d,0}$ and
+$w_{d,0}=w_{-d,0}$ (where all indices are taken modulo $N$, taking
+$-d\to N-d$).  To this end, we define a complete basis of orthogonal
+vectors that are symmetric under reflection, namely for all
+$d\in\VV_N\equiv\set{1,2,\cdots,\floor{N/2}}$,
+\begin{align}
+  \ket{d_+} \equiv \f{\eta_d}{\sqrt{2}}\p{\ket{d} + \ket{-d}},
+  && \eta_d \equiv
+  \begin{cases}
+    1 & d \ne N/2, \\
+    1/\sqrt{2} & d = N/2
+  \end{cases},
+\end{align}
+where the special case of $d=N/2$ affects only even $N$, for which
+$\ket{N/2}=\ket{-N/2}$.  We can use these vectors to define a
+projector onto the reflection-symmetric subspace and expand
+$\ul{\v{\m w}}$ as
+\begin{align}
+  \P_+ \equiv \sum_{d\in\VV_N} \op{d_+},
+  &&
+  \ul{\v{\m w}} = \sum_{d\in\VV_N} w_{d,0} \,
+  \eta_d \sqrt{2} \ket{d_+}.
+\end{align}
+Finally, we project $\ul{\check{\m h}}$ onto the reflection-symmetric
+subspace,
+\begin{align}
+  \P_+ \ul{\check{\m h}} \P_+
+  &= \sum_{a,b\in\VV_N} \sum_{c,d\ne0}
+  \p{h_{d,0} \delta_{cd} + h_{c-d,0}} \op{a_+} \op{d} \op{b_+} \\
+  &= \sum_{a,b\in\VV_N} \op{a_+}{b_+} \f12 \eta_a\eta_b
+  \sum_{c,d\ne0} \p{h_{d,0} \delta_{cd} + h_{c-d,0}}
+  \p{\delta_{ac} + \delta_{-a,c}} \p{\delta_{bd} + \delta_{-b,d}} \\
+  &= \sum_{a\in\VV_N} h_{a,0} \op{a_+}
+  + \sum_{a,b\in\VV_N} \eta_a\eta_b
+  \p{h_{a-b,0} + h_{a+b,0}} \op{a_+}{b_+},
+\end{align}
+which allows us to write down the reduced eigenvalue problem
+\begin{align}
+  \P_+ \ul{\check{\m h}} \P_+ \c \ul{\v{\m w}}
+  = \p{\f12\Delta + h_0} \ul{\v{\m w}}.
+\end{align}
+Note that the factors of $\eta_a,\eta_b$ in
+$\P_+ \ul{\check{\m h}} \P_+$ only matter when $N$ is even, in which
+case these factors merely multiply the last row and last column of
+$\P_+ \ul{\check{\m h}} \P_+$ by a factor of $1/\sqrt{2}$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Collective and multi-local operators in the fully symmetric

### 2019-12-12 soc_squeezing/collective_paper/review/response.tex
penultimate (I hope) changes in response to referee on collective spin paper

@@
   section of our manuscript (right column of page 3), changing the
-  sentence:
+  text:
 
@@
     considerations for the TST expansion requires additional analysis
-    that we defer to future work.}
+    that we defer to future work.  As we show from benchmarks of the
+    TST expansion in Section III, however, a detailed understanding of
+    these limitations is not necessary ...}
 
@@
     considerations for the TST expansion requires additional analysis
-    that we defer to future work, which may benefit from an
-    investigation of the connections between the TST expansion and
-    past works related to the Mori formalism[59, 60].}
+    that we defer to future work.  {\bf An investigation of
+      connections between the TST expansion and past work related to
+      the Mori formalism[59, 60], for example, might answer questions
+      about the breakdown and convergence of the TST expansion.}  As
+    we show from benchmarks of the TST expansion in Section III,
+    however, a detailed understanding of {\bf breakdown} is not
+    necessary ...}
 
@@
   as arbitrary coherent or incoherent combinations of these processes.
-  Similarly, we provide all ingredients necessary to account for
+  Similarly, we have done all of the work necessary to account for
   arbitrary combinations of {\it collective} decay, excitation, and
@@
   OAT, and do not consider such discussions relevant to our current
-  work.  Even with decoherence (of the types considered in our work),
-  the OAT model is an exactly analytically solvable (see
+  work.  Even with decoherence (of the types primarily considered in
+  our work), the OAT model is an exactly analytically solvable (see
   \url{https://doi.org/gcsjdp}).  Approximate numerical methods such
@@
     solution even in the presence of decoherence[64].  The approximate
-    and numerics-oriented TST expansion is therefore ill-suited for
-    studying the OAT model, which will merely serve as an exactly
-    solvable benchmark of our methods.  Wherever applicable, we will
-    provide exact results for the OAT model (see Appendix K, as well
-    as the Supplementary Material of Ref.~[14]).}
+    and numerics-oriented TST expansion is therefore an inappropriate
+    tool for studying the OAT model, which will merely serve as an
+    exactly solvable benchmark of our methods.  Wherever applicable,
+    we will provide exact results for the OAT model (see Appendix K,
+    as well as the Supplementary Material of Ref.~[14]).}
 
-  We hope that the referee and editors understand our position on this
-  matter -- namely, that we would rather avoid adding spurious
-  references to past works on a subject (the influence of decoherence
-  on spin squeezing in the exactly solvable OAT model) that is of
-  little relevance to our work on a numerical method to compute spin
-  correlators.
-
-\item Concerning the final question, about the robustness of the TNT
-  model:
+\item Concerning the final question, about the advantages of the TNT
+  model in the presence of decoherence:
 
@@
     comment on the reason?}
+
+  We thank the referee for their careful attention to our results.  We
+  agree that our findings warrant a comment on why TNT performs better
+  than OAT or TAT in the presence of decoherence.  We have therefore
+  rewritten most of the last paragraph of section III, changing:
+
+  \red{The results in Figure 2 show that TNT is much more robust to
+    single-spin decoherence than OAT or TAT, a finding that could not
+    necessarily be deduced from the weak-decoherence simulations
+    presented in Figure 1.  Strong-decoherence computations of this
+    sort ...}
+
+  to:
+
+  \green{The results in Figure 2 show that the TNT model can generate
+    more squeezing than the OAT or TAT models in the presence of
+    strong decoherence.  The better performance of TNT is a
+    consequence of the fact that TNT initially generates squeezing at
+    a faster rate than OAT or TAT, thereby allowing it to produce more
+    squeezing before the degrading effects of decoherence kick in.  We
+    corroborate this finding with quantum trajectory simulations of a
+    small system in Appendix L.  Strong-decoherence computations of
+    the sort used for Figure 2 ...}
+
+  The new appendix includes Monte Carlo simulations of a small system
+  with sufficiently strong decoherence to reproduce the improved
+  performance of TNT.  We hope that these additions address the
+  referee's questions and concerns about figure 2 to a satisfactory
+  degree.
+
+  % One consequence of applying our method in the large system, strong
+  % decoherence regime in figure 2 is that there is no way (barring a
+  % detailed analysis of error and breakdown in the TST expansion) to
+  % completely rule out numerical errors, as there are no other
+  % simulation methods available to corroborate the results.  We had
+  % hoped that the benchmarking in figure 1 and the discussion of
+  % breakdown mechanisms would assure readers that the TST expansion
+  % self-diagnoses its sudden departure from virtually exact results,
+  % thereby increasing confidence that the TST expansion can be
+  % trusted when no other methods are available.
 
@@
   in the new text).  We have therefore removed a subsequent definition
-  of this notation, which previously appeared in the following
-  paragraph, ``\red{where
-    $\abs{\v\ell}\equiv\sum_\alpha\ell_\alpha$}'' (middle right column
-  of page 2 in the old text).
+  of this notation, ``\red{where
+    $\abs{\v\ell}\equiv\sum_\alpha\ell_\alpha$}'', which previously
+  appeared in the following paragraph (middle right column of page 2
+  in the old text).
 

### 2019-12-12 soc_squeezing/collective_paper/collective_spin.tex
penultimate (I hope) changes in response to referee on collective spin paper

@@
 \begin{abstract}
-  We present a new method to compute short-time expectation values in
-  large collective spin systems with generic Markovian decoherence.
-  Our method is based on a Taylor expansion of a formal solution to
-  the equations of motion for Heisenberg operators.  This expansion
-  can be truncated at finite order to obtain virtually exact results
-  at short times that are relevant for metrological applications such
-  as spin squeezing.  In order to evaluate the expansion for
-  Heisenberg operators, we compute the relevant structure constants of
-  a collective spin operator algebra.  We demonstrate the utility of
-  our method by computing spin squeezing, two-time correlation
-  functions, and out-of-time-ordered correlators for $10^4$ spins in
-  strong-decoherence regimes that are otherwise inaccessible via
-  existing numerical methods.  Our method can be straightforwardly
-  generalized to the case of a collective spin coupled to bosonic
-  modes, relevant for trapped ion and cavity QED experiments, and may
-  be used to investigate short-time signatures of quantum chaos and
-  information scrambling.
+  We present an efficient method to compute short-time expectation
+  values in large collective spin systems with typical forms of
+  Markovian decoherence.  Our method is based on a Taylor expansion of
+  a formal solution to the equations of motion for Heisenberg
+  operators.  This expansion can be truncated at finite order to
+  obtain virtually exact results at short times that are relevant for
+  metrological applications such as spin squeezing.  In order to
+  evaluate the expansion for Heisenberg operators, we compute the
+  relevant structure constants of a collective spin operator algebra.
+  We demonstrate the utility of our method by computing spin
+  squeezing, two-time correlation functions, and out-of-time-ordered
+  correlators for $10^4$ spins in strong-decoherence regimes that are
+  otherwise inaccessible via existing numerical methods.  Our method
+  can be straightforwardly generalized to the case of a collective
+  spin coupled to bosonic modes, relevant for trapped ion and cavity
+  QED experiments, and may be used to investigate short-time
+  signatures of quantum chaos and information scrambling.
 \end{abstract}
@@
 genuinely multi-body correlations, however, eventually causes the
-cumulant expansion to yield unphysical results with no clear signature
+cumulant expansion to yield incorrect results with no clear signature
 of failure.  In the absence of other means to compute correlators, it
@@
 
-In this work, we present a new method to compute short-time dynamics
-of collective spin systems with generic Markovian decoherence.  The
-only restriction on decoherence (beyond Markovianity) is that, like
-the coherent collective dynamics, it must act identically on all
-constituent particles.  Our method is based on a short-time expansion
-of exact solutions to the equations of motion for Heisenberg
-operators.  Evaluating this expansion requires knowing the structure
+In this work, we present an efficient method to compute short-time
+dynamics of collective spin systems with typical forms of Markovian
+decoherence.  The only restriction on decoherence (beyond
+Markovianity) is that, like the coherent collective dynamics, it must
+act identically on all constituent particles.  Our method is based on
+a formal solution to the equations of motion for Heisenberg operators,
+thereby bearing some resemblance to the Mori
+formalism\cite{mori1965continuedfraction} and related
+work\cite{annett1994recursive}.  Specifically, we expand a formal
+solution for a Heisenberg operator into a Taylor series whose
+truncation can yield negligible error at sufficiently short times.
+Evaluating the resulting expansion requires knowing the structure
 constants of a collective spin operator algebra; the calculation of
@@
 
+We note that particle loss is an important decoherence mechanism in
+many experimental realizations of collective spin
+models\cite{ma2011quantum}.  In principle, a spin model has no notion
+of the particle annihilation operators that generate particle loss,
+and therefore cannot capture this effect directly.  Nonetheless, for a
+system initially composed of $N$ particles, the effect of particle
+loss can be emulated with $O(1/N)$ error by the dissipator
+$\check\D_{\t{loss}}$ defined by
+$\check\D_{\t{loss}}\hat\S_{\v m}=-\abs{\v m}\hat\S_{\v m}$, where
+$\abs{\v m}\equiv\sum_\alpha m_\alpha$ (see Appendix
+\ref{sec:particle_loss}).  Furthermore, the effect particle loss can
+be accounted for exactly by
+\begin{enumerate*}
+\item introducing an additional index on spin operators,
+  $\hat\S_{\v m}\to\hat\S_{N\v m}$, to keep track of different sectors
+  of fixed particle number within a multi-particle Fock space, and
+\item constructing jump operators that appropriately couple spin
+  operators within different particle-number sectors.
+\end{enumerate*}
+We defer a detailed exact accounting of particle loss to future work.
+
 The time derivative operator $\check T$ will generally couple spin
 operators $\hat\S_{\v n}$ to spin operators $\hat\S_{\v m}$ with
-higher ``weight'', i.e.~with $\abs{\v m}>\abs{\v n}$, where
-$\abs{\v\ell}\equiv\sum_\a\ell_\a$.  The growth of operator weight
-signifies the growth of many-body correlations.  Keeping track of this
-growth eventually becomes intractable, requiring us to truncate our
-equations of motion somehow.  The simplest truncation strategy would
-be to take
+higher ``weight'', i.e.~with $\abs{\v m}>\abs{\v n}$.  The growth of
+operator weight signifies the growth of many-body correlations.
+Keeping track of this growth eventually becomes intractable, requiring
+us to truncate our equations of motion somehow.  The simplest
+truncation strategy would be to take
 \begin{align}
@@
 considerations for the TST expansion requires additional analysis that
-we defer to future work.  As we show from benchmarks of the TST
-expansion in Section \ref{sec:squeezing}, however, a detailed
-understanding of these limitations is not necessary to diagnose the
-breakdown time $t_{\t{break}}^{(M)}$ beyond which the TST expansion
-yields inaccurate results.  Empirically, we find that going beyond
-order $M\approx35$ yields no significant gains in all cases considered
-in this work.
+we defer to future work.  An investigation of connections between the
+TST expansion and past work related to the Mori
+formalism\cite{mori1965continuedfraction, annett1994recursive}, for
+example, might answer questions about the breakdown and convergence of
+the TST expansion.  As we show from benchmarks of the TST expansion in
+Section \ref{sec:squeezing}, however, a detailed understanding of
+breakdown is not necessary to diagnose the breakdown time
+$t_{\t{break}}^{(M)}$ beyond which the TST expansion yields inaccurate
+results.  Empirically, we find that going beyond order $M\approx35$
+yields no significant gains in all cases considered in this work.
 
@@
 model, whose quantum dynamics admits an exact analytic solution even
-in the presence of decoherence\cite{foss-feig2013nonequilibrium}.  We
-will benchmark our calculations using these analytical OAT results
-wherever applicable (see Appendix \ref{sec:OAT}, as well as the
-Supplementary Material of Ref.~[\citenum{bohnet2016quantum}]).
+in the presence of decoherence\cite{foss-feig2013nonequilibrium}.  The
+approximate and numerics-oriented TST expansion is therefore an
+inappropriate tool for studying the OAT model, which will merely serve
+as an exactly solvable benchmark of our methods.  Wherever applicable,
+we will provide exact results for the OAT model (see Appendix
+\ref{sec:OAT}, as well as the Supplementary Material of
+Ref.~[\citenum{bohnet2016quantum}]).
 
@@
 drastic departure that can be diagnosed by inspection.  The breakdown
-of the TST expansion in Figure \ref{fig:benchmarking} is marked by an
-unphysical squeezing parameter $\xi^2<0$, which occurs due to
-individual correlators taking unphysical values with
-$\abs{\bk{\hat\S_{\v n}\p{t}}}\gtrsim S^{\abs{\v n}}$.  Going up
-through order $M=70$ in the TST expansion does not significantly
-increase the breakdown time $t_{\t{break}}^{(M)}$, and in some cases
-even shortens $t_{\t{break}}^{(M)}$.  The sudden and drastic departure
-from virtually exact results is consistent with the limitations of the
-TST expansion discussed at the end of Section \ref{sec:theory}.
-Specifically, we identify three possible mechanisms for breakdown:
+of the TST expansion in Figure \ref{fig:benchmarking} induces an
+unphysical squeezing parameter $\xi^2<0$.  In general, however, there
+is no fundamental relationship between the breakdown of the TST
+expansion and the conditions for a physical squeezing parameter
+$\xi^2$.  A proper diagnosis of breakdown therefore requires
+inspection of the correlators $\bk{\hat\S_{\v n}\p{t}}$ used to
+compute the squeezing parameter $\xi^2$, which upon breakdown will
+rapidly take unphysical values with
+$\abs{\bk{\hat\S_{\v n}\p{t}}}\gtrsim S^{\abs{\v n}}$.  The sudden and
+drastic departure from virtually exact results is consistent with the
+limitations of the TST expansion discussed at the end of Section
+\ref{sec:theory}.  Specifically, we identify three possible mechanisms
+for breakdown:
 \begin{enumerate*}
@@
 in parameter regimes that are inaccessible via other means to compute
-correlators.
+correlators.  Note that, due to the breakdown mechanisms of the TST
+expansion, going up through order $M=70$ does not significantly
+increase the breakdown time $t_{\t{break}}^{(M)}$ in Figure
+\ref{fig:benchmarking}, and in some cases even shortens
+$t_{\t{break}}^{(M)}$.
 
@@
 
-The results in Figure \ref{fig:decoherence_strong} show that TNT is
-much more robust to single-spin decoherence than OAT or TAT, a finding
-that could not necessarily be deduced from the weak-decoherence
-simulations presented in Figure \ref{fig:benchmarking}.
-Strong-decoherence computations of this sort were used to put lower
-bounds on theoretically achievable spin squeezing via TAT with
-decoherence in Ref.~[\citenum{he2019engineering}], exemplifying a
-concrete and practical application of the TST expansion and the
-collective-spin structure constants calculated in this work.
+The results in Figure \ref{fig:decoherence_strong} show that the TNT
+model can generate more squeezing than the OAT or TAT models in the
+presence of strong decoherence.  The better performance of TNT is a
+consequence of the fact that TNT initially generates squeezing at a
+faster rate than OAT or TAT, thereby allowing it to produce more
+squeezing before the degrading effects of decoherence kick in.  We
+corroborate this finding with quantum trajectory simulations of a
+small system in Appendix \ref{sec:small_system}.  Strong-decoherence
+computations of the sort used for Figure \ref{fig:decoherence_strong}
+were used to put lower bounds on theoretically achievable spin
+squeezing via TAT with decoherence in
+Ref.~[\citenum{he2019engineering}], exemplifying a concrete and
+practical application of the TST expansion and the collective-spin
+structure constants calculated in this work.
 
@@
 
-We have presented a new method for computing correlators at short
-times in collective spin systems.  This method is based on truncating
-a short-time expansion of Heisenberg operators, and can access
-correlators on time scales that are relevant to metrological
+We have presented an efficient method for computing correlators at
+short times in collective spin systems.  This method is based on
+truncating a short-time expansion of Heisenberg operators, and can
+access correlators on time scales that are relevant to metrological
 applications such as spin squeezing.  In order to evaluate the
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Emulating particle loss in a spin model}
+\label{sec:particle_loss}
+
+Here we discuss the details of emulating particle loss with $O(1/N)$
+error, where $N$ is the initial number of particles in a system that
+we wish to describe with a spin model.  Starting with the full algebra
+of creation and annihilation operators (whether bosonic or fermionic)
+in a system, spin models are typically implemented by identifying a
+subalgebra of relevant ``spin'' operators that satisfy appropriate
+commutation relations.  Two-state particles on a lattice, for example,
+are described by annihilation operators $c_{js}$ indexed by a lattice
+site $j\in\mathbb{Z}$ and an internal state index $s\in\set{\up,\dn}$,
+enabling the straightforward construction of spin operators
+\begin{align}
+  \sigma_\x^{(j)} \equiv c_{j,\up}^\dag c_{j,\dn} + \t{h.c.},
+  &&
+  \sigma_\y^{(j)} \equiv - i c_{j,\up}^\dag c_{j,\dn} + \t{h.c.},
+  &&
+  \sigma_\z^{(j)}
+  \equiv c_{j,\up}^\dag c_{j,\up} - c_{j,\dn}^\dag c_{j,\dn},
+  &&
+  \1^{(j)} \equiv c_{j,\up}^\dag c_{j,\up} + c_{j,\dn}^\dag c_{j,\dn},
+\end{align}
+which satisfy same commutation relations as the standard Pauli
+operators.  These spin operators can be more compactly defined in the
+form
+\begin{align}
+  \sigma_\alpha^{(j)} \equiv \sum_{r,s\in\set{\up,\dn}}
+  c_{jr}^\dag \bk{r|\sigma_\alpha|s} c_{js},
+\end{align}
+where $\sigma_\alpha$ for $\alpha\in\set{\x,\y,\z,\1}$ is a Pauli
+operator, with $\sigma_\1\equiv\1$; and $\bk{r|\sigma_\alpha|s}$
+denotes a matrix element of $\sigma_\alpha$.  This construction
+exemplifies how the set of jump operators
+$\J_{\t{loss}}^{\t{bare}}\equiv\set{c_{js}}$ that generate particle
+loss cannot be constructed from spin operators, which are generally
+bilinear in particle creation/annihilation operators.  When working on
+the level of a spin model, therefore, we can at best only {\it
+  emulate} the effect of particle loss by some indirect means.
+
+To understand the effect of particle loss on collective spin
+operators, we first define a single multi-body spin operator
+addressing sites $\v j=\set{j_1,j_2,\cdots,j_{\abs{\v j}}}$,
+\begin{align}
+  \sigma_{\v j\v\alpha} \equiv \prod_{j\in\v j} \sigma_{\alpha_j}^{(j)},
+\end{align}
+and expand
+\begin{align}
+  \D\p{\J_{\t{loss}}^{\t{bare}}} \sigma_{\v j\v\alpha}
+  &= \sum_{k,s} \p{c_{ks}^\dag \sigma_{\v j\v\alpha} c_{ks}
+    - \f12\sp{c_{ks}^\dag c_{ks}, \sigma_{\v j\v\alpha}}_+} \\
+  &= \sum_{k\in\v j} \sum_s c_{ks}^\dag \sigma_{\alpha_k}^{(k)} c_{ks}
+  \prod_{\substack{j\in\v j\\j\ne k}} \sigma_{\alpha_j}^{(j)}
+  + \sum_{k\notin\v j} \sum_s c_{ks}^\dag c_{ks} \sigma_{\v j\v\alpha}
+  - \f12 \sum_k \sp{\1^{(k)}, \sigma_{\v j\v\alpha}}_+ \\
+  &= \sum_{k\in\v j} \sum_{q,r,s} \bk{q|\sigma_{\alpha_k}|r}
+  c_{ks}^\dag c_{kq}^\dag c_{kr} c_{ks}
+  \prod_{\substack{j\in\v j\\j\ne k}} \sigma_{\alpha_j}^{(j)}
+  - \abs{\v j} \sigma_{\v j\v\alpha}.
+\end{align}
+In order to have an actual spin model, fermionic statistics or
+energetic considerations must forbid multiple occupation of individual
+lattice sites.  In that case, the on-site four-point product
+$c_{ks}^\dag c_{kq}^\dag c_{kr} c_{ks}=0$ vanishes, and
+\begin{align}
+  \D\p{\J_{\t{loss}}^{\t{bare}}} \sigma_{\v j\v\alpha}
+  = - \abs{\v j} \sigma_{\v j\v\alpha}.
+\end{align}
+Up to $O(1/N)$ corrections, a collective spin operator $\S_{\v m}$
+essentially consists of $\abs{\v m}$-body operators of the form
+$\sigma_{\v j\v\alpha}$ with $\abs{\v j}=\abs{\v m}$, which implies
+that the dissipator $\D_{\t{loss}}$ defined by
+$\D_{\t{loss}}\S_{\v m}=-\abs{\v m}\S_{\v m}$ describes particle loss
+with $O(1/N)$ error.  We note that the dissipator $\D_{\t{loss}}$ is
+essentially the depolarizing channel,
+i.e.~$\D_{\t{loss}}=\D\p{\J_{\t{loss}}}$ for
+$\J_{\t{loss}} = \set{s_\alpha^{(j)}}$ with $\alpha\in\set{\x,\y,\z}$
+and $j\in\set{1,2,\cdots,N}$.  A direct implementation of
+$\D_{\t{loss}}$ with $\D_{\t{loss}}\S_{\v m}=-\abs{\v m}\S_{\v m}$,
+however, is much more efficient than evaluating the depolarizing
+channel $\D\p{\J_{\t{loss}}}$ with the ingredients in Appendices
+\ref{sec:sandwich_single} and \ref{sec:decoherence_single}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Initial conditions}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Robustness of squeezing generated by the twist-and-turn (TNT)
+  model}
+\label{sec:small_system}
+
+% \begin{figure}
+%   \centering
+%   \includegraphics{decoherence_strong_small.pdf}
+%   \caption{Spin squeezing of $N=100$ spins initially in $\ket\X$ with
+%     spontaneous decay, excitation, and dephasing of individual spins
+%     at rates $\gamma_-=\gamma_+=\gamma_\z=2\chi$.  Computed computed
+%     using exact methods (solid line), quantum trajectory simulations
+%     with $10^4$ trajectories (dots), and the TST expansion in
+%     Eq.~\eqref{eq:TST} with $M=35$ (dashed lines).}
+%   \label{fig:decoherence_strong_small}
+% \end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Heisenberg operators in open quantum systems}
@@
 time-dependent coefficients $\O_{\v m}\p{t}$ for time-independent
-Schr\"odinger operators $\hat\S_{\v m}$.  Our discussion should
-clarify why noise operators play no role in our calculation of
-expectation values of the form $\bk{\O\p{t}}$ and $\bk{\O\Q\p{t}}$,
-despite the fact that noise operators generally {\it do} need to be
-considered in the calculation of more general multi-time correlators
-in open quantum systems\cite{blocher2019quantum}.
+Schr\"odinger operators $\S_{\v m}$.  Our discussion should clarify
+why noise operators play no role in our calculation of expectation
+values of the form $\bk{\O\p{t}}$ and $\bk{\O\Q\p{t}}$, despite the
+fact that noise operators generally {\it do} need to be considered in
+the calculation of more general multi-time correlators in open quantum
+systems\cite{blocher2019quantum}.
 
@@
   \U_{\ul S}\p{t} \rho_S
-  = \sum_j \U_S^{(j)}\p{t} \rho_S {\U_S^{(j)}}^\dag\p{t}
+  = \sum_j \U_S^{(j)}\p{t} \rho_S \U_S^{(j)\dag}\p{t}
 \end{align}
@@
   \equiv \U_{\ul S}^\dag\p{t} \O_S
-  = \sum_j {\U_S^{(j)}}^\dag\p{t} \O_S \U_S^{(j)}\p{t}.
+  = \sum_j \U_S^{(j)\dag}\p{t} \O_S \U_S^{(j)}\p{t}.
   \label{eq:heisenberg_wrong}

### 2020-01-13 collective_sun/sun_perturbation.tex
discuss the point of partitions in Appendix D, improve a bit of notation around e.g. v vs v_\Delta

@@
 $w_{pq}=w_{qp}$ and $w_{pp}=0$.
+% TODO: consider more general perturbations, namely:
+% (1) a sum over X in V_1,
+% (2) more general two-body perturbations (O_pq) in V_2, and
+% (3) a sum over different O_pq in V_2
 
@@
 states by an interaction energy difference that is large compared to
-any coupling between $\M_0$ and its orthogonal complement.
+any coupling between $\M_0$ and its orthogonal complement.  The
+precise conditions for such a non-vanishing many-body interaction
+energy gap is an important question that deserves further study in
+future work.
 
@@
   &&
-  \E \equiv \sum_{\Delta>0} \f{\P_\Delta}{\Delta},
+  \E \equiv \sum_{\Delta\ne0} \f{\P_\Delta}{\Delta},
 \end{align}
@@
 \begin{align}
-  H_0 \V_1 \ket\psi
-  = E_0 \V_1 \ket\psi
+  H_0 \V_1\p{\v v} \ket\psi
+  = E_0 \V_1\p{\v v} \ket\psi
   + \p{\sum_{p,q} h_{pq} v_p X_q - \sum_q h_q v_q X_q} \ket\psi,
+  &&
+  h_q \equiv \sum_p h_{pq}.
   \label{eq:H_V1_psi}
 \end{align}
-where $h_q\equiv\sum_p h_{pq}$ is the net interaction energy
-associated with spin $p$.  We thus find that the vector $\V_1\ket\psi$
-is an eigenvector of the interaction Hamiltonian $H_0$ if the
-coefficients $\v v$ satisfy
-\begin{align}
-  \p{\m h - \diag\v h}\c\v v = \Delta\p{\v v} \v v,
+We thus find that $\V_1\p{\v v_\Delta}\ket\psi$ is an eigenvector of
+the interaction Hamiltonian $H_0$ if the coefficients $\v v_\Delta$
+satisfy
+\begin{align}
+  \p{\m h - \diag\v h}\c\v v_\Delta = \Delta \v v_\Delta,
   \label{eq:cond_1}
@@
 $h_p,h_{pq}$; $\diag\v h$ is a matrix with $\v h$ on the diagonal and
-zeros everywhere else; and $\Delta\p{\v v}$ is some constant that may
-depend on $\v v$.  Solving the general eigenvalue problem in
-\eqref{eq:cond_1} requires diagonalizing an $N\times N$ matrix.  When
-the interaction Hamiltonian $H_0$ is translationally invariant,
-however, this eigenvalue problem can be solved analytically; we
-provide the corresponding solution in Appendix \ref{sec:trans_inv}.
+zeros everywhere else; and $\Delta$ is a scalar.  Solving the general
+eigenvalue problem in \eqref{eq:cond_1} requires diagonalizing an
+$N\times N$ matrix.  When the interaction Hamiltonian $H_0$ is
+translationally invariant, however, this eigenvalue problem can be
+solved analytically; we provide the corresponding solution in Appendix
+\ref{sec:trans_inv}.
 
@@
 $\v v_\Delta$, and is entirely independent of the state $\ket\psi$ or
-choice of single-spin trace-zero operator $X$ used to build $\V_1$ in
-\eqref{eq:perturbations}.  Finding operators $\V_1\p{\v v_\Delta}$
-that generate eigenvectors of the interaction Hamiltonian $H_0$ when
-they act on fully symmetric states $\ket\psi$ thus reduces to finding
-eigenvectors $\v v_\Delta$ of the interaction matrix $\m h-\diag\v h$.
+the choice of a single-spin trace-zero operator $X$ used to build
+$\V_1$ in \eqref{eq:perturbations}.
+% TODO: consider more general single-body perturbations V_1, which sum
+% over different operators X
+Finding operators $\V_1\p{\v v_\Delta}$ that generate eigenvectors of
+the interaction Hamiltonian $H_0$ when they act on fully symmetric
+states $\ket\psi$ thus reduces to finding eigenvectors $\v v_\Delta$
+of the interaction matrix $\m h-\diag\v h$.
 
@@
   = - \P_0 \V_1\p{\v v} \E \V_1\p{\v v} \P_0
-  = - \sum_{\Delta>0} \f1{\Delta}
+  = - \sum_{\Delta\ne0} \f1{\Delta}
   \P_0 \V_1\p{\v v} \P_\Delta \V_1\p{\v v} \P_0
-  = - \sum_{\Delta>0} \f1{\Delta} \P_0 \V_1\p{\v v_\Delta}^2 \P_0,
+  = - \sum_{\Delta\ne0} \f1{\Delta} \P_0 \V_1\p{\v v_\Delta}^2 \P_0,
 \end{align}
@@
   \EQFS \f1{N\p{N-1}}
-  \sum_{\Delta>0} \f{\v v_\Delta\c\v v_\Delta}{\Delta}
-  \sp{\p{\col{X}}^2 - N \p{\col{X^2}}},
+  \sum_{\Delta\ne0} \f{\v v_\Delta\c\v v_\Delta}{\Delta}
+  \sp{{\col{X}\,}^2 - N \col{X^2}},
   \label{eq:H_1_2}
@@
 \begin{align}
-  H_0 \V_2 \ket\psi
-  = E_0 \V_2 \ket\psi
+  H_0 \V_2\p{\m w} \ket\psi
+  = E_0 \V_2\p{\m w} \ket\psi
   + \f12 \sum_{p\ne q} \sp{2\bar h_{pq} w_{pq}
@@
 \end{align}
-We thus find that the vector $\V_2\ket\psi$ is an eigenvector of the
-interaction Hamiltonian $H_0$ if the coefficients $\m w$ satisfy
-\begin{align}
-  2 \bar h_{pq} w_{pq}
-  + \sum_{k\notin\set{p,q}} \p{h_{pk} w_{kq} + w_{pk} h_{kq}}
-  = \Delta\p{\m w} w_{pq},
+We thus find that the vector $\V_2\p{\m w_\Delta}\ket\psi$ is an
+eigenvector of the interaction Hamiltonian $H_0$ if the coefficients
+$\m w_\Delta$ satisfy the eigenvalue equation
+\begin{align}
+  \check{\m h} \c \v{\m w}_\Delta = \Delta \v{\m w}_\Delta,
   \label{eq:cond_2}
 \end{align}
-where $\Delta\p{\m w}$ is some constant that may depend on $\m w$, and
-for clarity we explicitly exclude terms with $k\in\set{p,q}$ that
-vanish due to the fact that $\m h$ and $\m w$ are zero on the
-diagonal.  We can recast the condition in \eqref{eq:cond_2} as an
-eigenvalue problem by writing the coefficient matrix $\m w$ as a
-vector $\v{\m w}\in\RR^{N\choose 2}$,
-\begin{align}
-  \m w = \sum_{p,q} w_{pq} \op{p}{q}
-  \to \v{\m w} \equiv \flat\m w
-  \equiv \sum_{p<q} w_{pq} \ket{\set{p,q}},
-\end{align}
-where we label basis of vectors $\ket{\set{p,q}}\in\RR^{N\choose 2}$
-by a choice of two distinct integers $p,q\in\ZZ_N$.  We similarly
-define the ${N \choose 2}\times{N \choose 2}$ matrix
+where for any symmetric $N\times N$ matrix $\m m$ with zeroes on the
+diagonal (as is the case with the coefficient matrix $\m w_\Delta$),
+we define the ${N \choose 2}$-component vector
+\begin{align}
+  \v{\m m} \equiv \sum_{p<q} m_{pq} \ket{\set{p,q}},
+  \label{eq:flat_vec}
+\end{align}
+and we define the ${N \choose 2} \times {N \choose 2}$ matrix
 \begin{align}
@@
 \end{align}
-The condition in \eqref{eq:cond_2} is then equivalently the eigenvalue
-equation
-\begin{align}
-  \check{\m h} \c \v{\m w} = \Delta\p{\m w} \v{\m w}.
-  \label{eq:cond_2_eig}
-\end{align}
-When the interaction Hamiltonian $H_0$ and perturbation $\V_2\p{\m w}$
-are translationally invariant, this eigenvalue problem can be reduced
-to that of diagonalizing a matrix of dimensions
-$\floor{N/2}\times\floor{N/2}$, as we show in Appendix
+The basis vectors $\ket{\set{p,q}}$ are labeled by a choice
+$\set{p,q}$ of two distinct spins $p,q$.  The eigenvalue problem in
+\eqref{eq:cond_2} can be simplified when the interaction Hamiltonian
+$H_0$ and the perturbation $\V_2\p{\m w}$ are translationally
+invariant; we discuss this simplification in Appendix
 \ref{sec:trans_inv}.
@@
 If we {\it construct} a perturbation $\V_2\p{\m w_\Delta}$ with
-coefficients $\m w_\Delta$ that satisfy \eqref{eq:cond_2_eig} with
+coefficients $\m w_\Delta$ that satisfy \eqref{eq:cond_2} with
 eigenvalue $\Delta$, then
@@
 matrix $\m w$ into its projections $\m w_\Delta$ onto the eigenspace
-of matrices satisfying \eqref{eq:cond_2_eig} with eigenvalue $\Delta$,
+of matrices satisfying \eqref{eq:cond_2} with eigenvalue $\Delta$,
 i.e.~$\m w=\sum_\Delta\m w_\Delta$, which allows us to expand
@@
   H_2^{(2)} = - \P_0 \V_2\p{\m w} \E \V_2\p{\m w} \P_0
-  = -\sum_{\Delta>0} \f1\Delta \P_0 \V_2\p{\m w_\Delta}^2 \P_0,
+  = -\sum_{\Delta\ne0} \f1\Delta \P_0 \V_2\p{\m w_\Delta}^2 \P_0,
 \end{align}
@@
 $X=Y=s_\z$, where $s_\z\equiv\p{\op\up-\op\dn}/2$ is an SU(2) spin
-operator, then we can expand
+operator, then
 \begin{align}
@@
 In order to expand the coefficients $W_k\p{\m h,\m w}$, for any
-symmetric matrix $\m m$ with zeroes on the diagonal we define the
-vectors
-\begin{align}
-  \v{\m m} \equiv \flat\m m \equiv \sum_{p<q} m_{pq} \ket{\set{p,q}},
-  &&
+symmetric matrix $\m m$ with zeros on the diagonal, similarly to the
+special definition $\v{\m m}$ in \eqref{eq:flat_vec} we define
+\begin{align}
   \v m \equiv \sum_{p,q} m_{pq} \ket{p},
 \end{align}
-in terms of which
+which is essentially the sum over all columns (or rows) of $\m m$, in
+terms of which
 \begin{align}
   W_4\p{\m h,\m w}
-  \equiv \sum_\Delta \f1\Delta \f{\v w_\Delta\c\v w_\Delta
+  \equiv \sum_{\Delta\ne0} \f1\Delta \f{\v w_\Delta\c\v w_\Delta
     - \v{\m w}_\Delta\c\v{\m w}_\Delta}{N!_4},
@@
   W_2\p{\m h,\m w}
-  \equiv \sum_\Delta \f1{\Delta} \sp{\f{\v w_\Delta\c\v w_\Delta
+  \equiv \sum_{\Delta\ne0} \f1{\Delta} \sp{\f{\v w_\Delta\c\v w_\Delta
       - 2\v{\m w}_\Delta\c\v{\m w}_\Delta}{4N!_2}
@@
 \end{align}
-Note that the dependence of the coefficients of $W_k\p{\m h,\m w}$ on
-the interaction matrix $\m h$ is implicit in the decomposition of the
+The dependence of the coefficients of $W_k\p{\m h,\m w}$ on the
+interaction matrix $\m h$ is implicit in the decomposition of the
 coefficient matrix $\m w$ into components $\m w_\Delta$ as
 $\m w=\sum_\Delta\m w_\Delta$, where $\m w_\Delta$ is the projection
-of $\m w$ onto the $\Delta$-eigenspace of the operator $\check{\m h}$
+of $\m w$ onto the $\Delta$-eigenspace of the matrix $\check{\m h}$
 defined in \eqref{eq:h_super_mat}.
@@
 \begin{align}
-  H_0 \V_2 \ket\psi
+  H_0 \V_2\p{\m w} \ket\psi
   = \f12 \sum_{\substack{k,\ell\\p<q}} w_{k\ell} h_{pq}
@@
 \begin{align}
-  \p{\col{X}}^2
+  {\col{X}\,}^2
   = \sum_p X_p^2 + \sum_{p\ne q} X_p X_q
@@
   \EQFS -\f{\v v\c\v v}{N\p{N-1}}
-  \sp{\p{\col{X}}^2 - N \p{\col{X^2}}}
+  \sp{{\col{X}\,}^2 - \col{X^2}}
   \label{eq:PXXP_neq}
@@
   \EQFS -\f{\v v\c\v v}{N\p{N-1}}
-  \sp{\p{\col{X}}^2 - N \p{\col{X^2}}}.
+  \sp{{\col{X}\,}^2 - N \col{X^2}}.
 \end{align}
@@
 \begin{align}
-  \P_0 \V_2 \P_0
+  \P_0 \V_2\p{\m w} \P_0
   \EQFS \f12 \bar w \p{\col{X}\odot\col{Y} - \col{X\odot Y}}.
@@
 \begin{align}
-  \P_0 \V_2^2 \P_0
+  \P_0 \V_2\p{\m w}^2 \P_0
   \EQFS \f14 \sum_{k,\ell,p,q}
@@
 \begin{multline}
-  \P_0 \V_2^2 \P_0
+  \P_0 \V_2\p{\m w}^2 \P_0
   \EQFS -\p{\v w\c\v w - \v{\m w}\c\v{\m w}} X_0 X_1 Y_2 Y_3
@@
 \begin{align}
-  \p{\m h - \diag\v h}\c\v v = \Delta\p{\v v} \v v,
+  \p{\m h - \diag\v h}\c\v v = \Delta \v v,
   \label{eq:cond_1_ref}
@@
 
-The two-body eigenvalue equation in \eqref{eq:cond_2_eig} is designed
-to find solutions $\m w$ to the equation
+The two-body eigenvalue equation in \eqref{eq:cond_2} is designed to
+find solutions $\m w$ to the equation
 \begin{align}
@@
   \ket{d_+} \equiv \f{\eta_d}{\sqrt{2}}\p{\ket{d} + \ket{-d}},
-  && \eta_d \equiv
+  &&
+  \eta_d \equiv
   \begin{cases}
@@
 invariance under arbitrary permutations of the single-spin operators
-$O\in S$.  To this end, for any partition\footnote{A partition
-  $P\equiv\set{p}$ of $S$ is a set of mutually disjoint, non-empty
-  subsets $p\subset S$ with $\bigcup_{p\in P}p=S$.}  $P\equiv\set{p}$
-of $S$ we define the $\abs{P}$-local operator
+$O\in S$.
+
+Even though $S$ contains $\abs{S}$ spins, a collective operator built
+from the single-spin operators $O\in S$ may have multiple
+(single-spin) operators acting on the same spin, resulting in a
+$k$-local operator with $k<\abs{S}$.  We therefore {\it
+  partition}\footnote{A partition $P\equiv\set{p}$ of $S$ is a set of
+  mutually disjoint, non-empty subsets $p\subset S$ with
+  $\bigcup_{p\in P}p=S$.} $S$ into subsets of operators that act on
+the same spin.  For any partition $P\equiv\set{p}$ of $S$ (here
+$p\subset S$ is a subset of operators $O\in S$ that address the same
+spin), we then define the $\abs{P}$-local operator
 \begin{align}
@@
 \end{align}
-In words, a partition $P$ of $S$ consists of an assignment of
-operators $O\in S$ to individual parts $p\subset S$; each part $p$ is
-associated with a tensor factor of $O_P$, and the operators $O\in p$
-act on that tensor factor.  The set of multi-local operators
-$\set{O_P:\t{partitions}~P~\t{of}~S}$ consists of all operators that
+The set of multi-local operators $\set{O_P:\t{partitions}~P~\t{of}~S}$
+consists of all operators that
 \begin{enumerate*}
-\item can be built from ordinary and tensor products of all
+\item can be built from ordinary products and tensor products of all
   single-spin operators $O\in S$, and
-\item are invariant under arbitrary permutations of $S$ (after
-  projection onto the fully symmetric manifold).
+\item are invariant under arbitrary permutations of $S$.
 \end{enumerate*}
@@
 \end{align}
-where we underline a single-spin operator to denote its collective
-version, i.e.~$\col{Q}\equiv\sum_j Q^{(j)}$, where $Q^{(j)}$ is an
-operator that acts with the single-spin operator $Q$ on spin $j$ and
-trivially on all other spins.  By construction, the set of collective
-operators $\set{\O_P:\t{partitions}~P~\t{of}~S}$ is one-to-one to the
-set of multi-local operators $\set{O_P:\t{partitions}~P~\t{of}~S}$,
-and it is straightforward to work out that for any set $S$ of
-single-spin operators,
+where we place a bar over a single-spin operator to denote its
+collective version, i.e.~$\col{Q}\equiv\sum_j Q^{(j)}$, where
+$Q^{(j)}$ is an operator that acts with the single-spin operator $Q$
+on spin $j$ and trivially on all other spins.  By construction, the
+set of collective operators $\set{\O_P:\t{partitions}~P~\t{of}~S}$ is
+one-to-one to the set of multi-local operators
+$\set{O_P:\t{partitions}~P~\t{of}~S}$, and it is straightforward to
+work out that for any set $S$ of single-spin operators,
 \begin{align}
@@
 \begin{align}
-  N!_2\, X\otimes X \EQFS \p{\col{X}}^2 - \p{\col{X^2}},
-  &&
-  N!_3\, X^{\otimes 3} \EQFS \p{\col{X}}^3
-  - 3 \col{X} \odot \col{X^2} + 2\p{\col{X^3}},
+  N!_2\, X\otimes X \EQFS {\col{X}\,}^2 - \col{X^2},
+  &&
+  N!_3\, X^{\otimes 3} \EQFS {\col{X}\,}^3
+  - 3 \col{X} \odot \col{X^2} + 2\col{X^3},
 \end{align}
@@
   N!_4\, X^{\otimes 4}
-  \EQFS \p{\col{X}}^4 - 4 \p{\col{X}}^2\odot\p{\col{X^2}}
-  - 2 \col{X} \p{\col{X}}^2 \col{X}
-  + 3 \p{\col{X^2}}^2 + 8 \col{X} \odot \p{\col{X^3}}
-  - 6 \p{\col{X^4}}.
+  \EQFS {\col{X}\,}^4 - 4 {\col{X}\,}^2 \odot \col{X^2}
+  - 2 \col{X} \, {\col{X}\,}^2 \, \col{X}
+  + 3 {\col{X^2}\,}^2 + 8 \col{X} \odot \col{X^3}
+  - 6 \col{X^4}.
 \end{align}
@@
 \begin{align}
-  N!_2\, X\otimes X \EQFS \p{\col{X}}^2 - N,
-  &&
-  N!_3\, X^{\otimes 3} \EQFS \p{\col{X}}^3 - \p{3N-2} \col{X},
-\end{align}
-\begin{align}
-  N!_4\, X^{\otimes 4} \EQFS \p{\col{X}}^4
-  - 2\p{3N-4} \p{\col{X}}^2 + 3 N \p{N-2}.
+  N!_2\, X\otimes X \EQFS {\col{X}\,}^2 - N,
+  &&
+  N!_3\, X^{\otimes 3} \EQFS {\col{X}\,}^3 - \p{3N-2} \col{X},
+\end{align}
+\begin{align}
+  N!_4\, X^{\otimes 4} \EQFS {\col{X}\,}^4
+  - 2\p{3N-4} {\col{X}\,}^2 + 3 N \p{N-2}.
 \end{align}

### 2020-01-13 collective_sun/sun_perturbation.tex
work out the perturbation theory for more general single- and two-body perturbations

@@
 \begin{align}
-  H_0 = \sum_{p<q} h_{pq} \Pi_{pq}
-  = \f12 \sum_{p,q} h_{pq} \Pi_{pq},
+  H_0 = \sum_{p<q} h_{pq} \Pi_{pq},
   &&
@@
 \end{align}
-where the coefficients $h_{pq}$ in the Hamiltonian satisfy
-$h_{pq}=h_{qp}$ and $h_{pp}=0$;
+where $p,q$ index individual spins; $\mu,\nu$ index states in an
+orthonormal basis for the $n$-dimensional Hilbert space $\H_n$ of a
+single spin; $h_{pq}$ are scalar coefficients; the operator
 $S_{\mu\nu}^{(p)}\equiv\op{\mu}{\nu}_p$ flips the state of spin $p$ to
-$\ket\mu$ from $\ket\nu$; and the operator $\Pi_{pq}$ permutes spins
-$p$ and $q$.  The permutation operator $\Pi_{pq}$ is a multilevel
-generalization of the SU(2)-symmetric spin interaction
-$\m S_p\c\m S_q$.
+$\ket\mu$ from $\ket\nu$; and the operator $\Pi_{pq}$ permutes the
+states of spins $p$ and $q$.  The permutation operator $\Pi_{pq}$ is a
+multilevel generalization of the SU(2)-symmetric spin interaction
+$\v S_p\c\v S_q$ for $\v S=\p{S_\x,S_\y,S_\z}$.
 
@@
 \begin{align}
-  \V_1 \equiv \sum_p v_p X_p = \v v\c\v X,
-  &&
-  \V_2 \equiv \sum_{p<q} w_{pq} \times \f12 \p{X_p Y_q + X_q Y_p}
-  = \f12 \v X \c\m w \c\v Y,
+  \V_1 \equiv \sum_X \sum_p v_{Xp} X_p = \sum_X \v v_X\c\v X,
+  &&
+  \V_2 \equiv \sum_O \sum_{p<q} w_{Opq} O_{pq}
+  = \sum_O \v{\m w}_O\c \v{\m O}
   \label{eq:perturbations}
 \end{align}
-where $X,Y$ are self-adjoint\footnote{In future versions of these
-  notes, we should relax the restriction that $X,Y$ are self-adjoint.}
-trace-zero operators on an $n$-dimensional Hilbert space; $X_p,Y_p$
-denote the action of $X,Y$ on spin $p$; $\v X,$ $\v Y$, $\v v$, and
-$\m w$ are respectively vectors and a matrix of all
-$X_p,Y_p,v_p,w_{pq}$; we use a dot ($\c$) to denote an inner product
-between vectors and matrices as appropriate; and we enforce
-$w_{pq}=w_{qp}$ and $w_{pp}=0$.
-% TODO: consider more general perturbations, namely:
-% (1) a sum over X in V_1,
-% (2) more general two-body perturbations (O_pq) in V_2, and
-% (3) a sum over different O_pq in V_2
+where $p,q$ index individual spins; $v_{Xp}$ and $w_{Opq}$ are real
+numbers; $X$ is a trace-zero single-body operator on $\H_n$; $O$ is a
+trace-zero two-body operator on $\H_n\otimes\H_n$; $\v v_X$ and $\v X$
+are $N$-component vectors defined by
+\begin{align}
+  \v v_X \equiv \sum_p v_{Xp} \ket{p},
+  &&
+  \v X \equiv \sum_p X_p \ket{p};
+\end{align}
+and finally $\v{\m w_O}$ and $\v{\m O}$ are ${N \choose 2}$-component
+vectors defined by
+\begin{align}
+  \v{\m w}_O \equiv \sum_{p<q} w_{Opq} \ket{\set{p,q}},
+  &&
+  \v{\m O} \equiv \sum_{p<q} O_{pq} \ket{\set{p,q}},
+\end{align}
+where the basis vectors $\ket{\set{p,q}}$ are labeled by a choice of
+two distinct spins $p,q$.
 
@@
 \section{Single-body perturbations}
-\label{sec:single_body}
+\label{sec:single_body_pert}
 
@@
 \begin{align}
-  H_1^{(1)} = \sum_p v_p \P_0 X_p \P_0
-  = \sum_p v_p \P_0 X_0 \P_0 \EQFS \bar v \col{X},
+  H_1^{(1)} = \sum_{X,p} v_{Xp} \P_0 X_p \P_0
+  \EQFS \sum_{X,p} v_{Xp} X_0
+  \EQFS N \sum_X \bar v_X \col{X},
   \label{eq:H_1_1}
@@
 \begin{align}
-  \bar v \equiv \f1N \sum_p v_p,
-  &&
-  \col{X} \equiv \sum_p X_p.
+  \bar v_X \equiv \f1N \sum_p v_{Xp},
+  &&
+  \col{X} \equiv \f1N \sum_p X_p.
 \end{align}
@@
 \begin{align}
-  H_0 \V_1\p{\v v} \ket\psi
-  = E_0 \V_1\p{\v v} \ket\psi
-  + \p{\sum_{p,q} h_{pq} v_p X_q - \sum_q h_q v_q X_q} \ket\psi,
-  &&
-  h_q \equiv \sum_p h_{pq}.
+  H_0 \V_1 \ket\psi
+  = E_0 \V_1 \ket\psi
+  + \sum_X \p{\sum_{p,q} h_{pq} v_{Xp} X_q - \sum_p h_p v_{Xp} X_p}
+  \ket\psi,
+  &&
+  h_p \equiv \sum_q h_{pq},
   \label{eq:H_V1_psi}
 \end{align}
-We thus find that $\V_1\p{\v v_\Delta}\ket\psi$ is an eigenvector of
-the interaction Hamiltonian $H_0$ if the coefficients $\v v_\Delta$
-satisfy
-\begin{align}
-  \p{\m h - \diag\v h}\c\v v_\Delta = \Delta \v v_\Delta,
+where we define $h_{pq}$ for all $p,q$ (as opposed to only $p<q$) by
+enforcing $h_{pq}=h_{qp}$ and $h_{pp}=0$.  We thus find that a
+perturbation $\V_1^\Delta\ket\psi$ indexed by a scalar $\Delta$ is an
+eigenvector of the interaction Hamiltonian $H_0$ if its coefficient
+vectors $\v v_X^\Delta$ all satisfy the eigenvalue equation
+\begin{align}
+  \p{\m h - \diag\v h}\c\v v = \Delta \v v,
+  &&
+  \m h \equiv \sum_{p,q} h_{pq} \op{p}{q},
+  &&
+  \v h \equiv \sum_q h_q \ket{q},
   \label{eq:cond_1}
 \end{align}
-where $\v h,\m h$ are respectively a vector and matrix of all
-$h_p,h_{pq}$; $\diag\v h$ is a matrix with $\v h$ on the diagonal and
-zeros everywhere else; and $\Delta$ is a scalar.  Solving the general
-eigenvalue problem in \eqref{eq:cond_1} requires diagonalizing an
-$N\times N$ matrix.  When the interaction Hamiltonian $H_0$ is
-translationally invariant, however, this eigenvalue problem can be
-solved analytically; we provide the corresponding solution in Appendix
-\ref{sec:trans_inv}.
-
-If we {\it construct} a perturbation $\V_1\p{\v v_\Delta}$ with
-coefficients $\v v_\Delta$ that satisfy \eqref{eq:cond_1} with
-eigenvalue $\Delta$, then
-\begin{align}
-  H_0 \V_1\p{\v v_\Delta} \ket\psi
-  = \p{E_0 + \Delta} \V_1\p{\v v_\Delta} \ket\psi.
+where $\diag\v h\equiv\sum_p h_p \op{p}$ is a matrix with $\v h$ on
+the diagonal and zeros everywhere else.  Solving the general
+eigenvalue problem in \eqref{eq:cond_1} requires diagonalizing the
+$N\times N$ matrix $\m h-\diag\v h$.  When the interaction Hamiltonian
+$H_0$ is translationally invariant, however, this eigenvalue problem
+can be solved analytically; we provide the corresponding solution in
+Appendix \ref{sec:trans_inv}.
+
+If we {\it construct} a perturbation $\V_1^\Delta$ with coefficients
+$\v v_X^\Delta$ that satisfy \eqref{eq:cond_1} with eigenvalue
+$\Delta$, then
+\begin{align}
+  H_0 \V_1^\Delta \ket\psi = \p{E_0 + \Delta} \V_1^\Delta \ket\psi.
 \end{align}
 Interestingly, the energy of the (unnormalized) state
-$\V_1\p{\v v_\Delta}\ket\psi$ depends only on the coefficients
-$\v v_\Delta$, and is entirely independent of the state $\ket\psi$ or
-the choice of a single-spin trace-zero operator $X$ used to build
-$\V_1$ in \eqref{eq:perturbations}.
-% TODO: consider more general single-body perturbations V_1, which sum
-% over different operators X
-Finding operators $\V_1\p{\v v_\Delta}$ that generate eigenvectors of
-the interaction Hamiltonian $H_0$ when they act on fully symmetric
-states $\ket\psi$ thus reduces to finding eigenvectors $\v v_\Delta$
+$\V_1^\Delta\ket\psi$ depends only on the coefficients
+$\v v_X^\Delta$, and is entirely independent of the state $\ket\psi$
+or the choice of a trace-zero single-body operators $X$ used to build
+$\V_1^\Delta$.  Finding operators $\V_1^\Delta$ that generate
+eigenvectors of the interaction Hamiltonian $H_0$ when they act on
+fully symmetric states $\ket\psi$ thus reduces to finding eigenvectors
 of the interaction matrix $\m h-\diag\v h$.
@@
 If the fully symmetric manifold $\M_0$ is gapped away from all other
-states, then then a vector $\V_1\p{\v v_\Delta}\ket\psi$ with
-$\Delta=0$ must lie within the fully symmetric manifold $\M_0$, which
-implies that the operator $\V_1\p{\v v_\Delta}$ preserves the
-permutational symmetry of $\M_0$.  Indeed, a constant vector
-$\v\1\equiv\p{1,1,1,\cdots}$ of ones satisfies the condition in
-\eqref{eq:cond_1} with eigenvalue $0$.  All other vectors
-$\v v_\Delta$ satisfying \eqref{eq:cond_1} with $\Delta\ne0$ must be
-orthogonal to $\v\1$, and therefore mean-zero.
+states, then then a vector $\V_1^\Delta\ket\psi$ with $\Delta=0$ must
+lie within the fully symmetric manifold $\M_0$, which implies that the
+operator $\V_1^\Delta$ preserves the permutational symmetry of $\M_0$.
+Indeed, a constant vector $\v\1\equiv\p{1,1,1,\cdots}$ of ones
+satisfies the condition in \eqref{eq:cond_1} with eigenvalue $0$,
+which implies that all vectors $\v v$ satisfying \eqref{eq:cond_1}
+with $\Delta\ne0$ must be orthogonal to $\v\1$, and therefore
+mean-zero.
 
 We now return to the task of computing the second-order effective
-Hamiltonian $H_1^{(2)}$.  Any coefficient vector $\v v$ can be
-expanded into its projections $\v v_\Delta$ onto the eigenspace of
+Hamiltonian $H_1^{(2)}$.  Any coefficient vector $\v v_X$ can be
+expanded into its projections $\v v_X^\Delta$ onto the eigenspace of
 vectors satisfying \eqref{eq:cond_1} with eigenvalue $\Delta$,
-i.e.~$\v v = \sum_\Delta \v v_\Delta$, which also allows us to expand
-\begin{align}
-  \V_1\p{\v v} = \sum_\Delta \V_1\p{\v v_\Delta},
-\end{align}
-where each operator $\V_1\p{\v v_\Delta}$ generates a state with
-definite interaction energy $\Delta$ above that of the fully symmetric
-manifold $\M_0$.  We can therefore simplify
+i.e.~$\v v_X = \sum_\Delta \v v_X^\Delta$, which also allows us to
+expand
+\begin{align}
+  \V_1 = \sum_\Delta \V_1^\Delta,
+  &&
+  \V_1^\Delta \equiv \sum_X \v v_X^\Delta \c \v X,
+\end{align}
+where each operator $\V_1^\Delta$ generates a state with definite
+interaction energy $\Delta$ above that of the fully symmetric manifold
+$\M_0$.  We can therefore simplify
 \begin{align}
   H_1^{(2)}
-  = - \P_0 \V_1\p{\v v} \E \V_1\p{\v v} \P_0
+  = - \P_0 \V_1 \E \V_1 \P_0
   = - \sum_{\Delta\ne0} \f1{\Delta}
-  \P_0 \V_1\p{\v v} \P_\Delta \V_1\p{\v v} \P_0
-  = - \sum_{\Delta\ne0} \f1{\Delta} \P_0 \V_1\p{\v v_\Delta}^2 \P_0,
-\end{align}
-where the product $\P_0 \V_1\p{\v v_\Delta}^2 \P_0$ is worked out in
-Appendix \ref{sec:PXXP}.  In total, we find
+  \P_0 \V_1^\Delta \P_\Delta \V_1^\Delta \P_0
+  = - \sum_{\Delta\ne0} \f1{\Delta} \P_0 \p{\V_1^\Delta}^2 \P_0,
+\end{align}
+where the product $\P_0 \p{\V_1^\Delta}^2 \P_0$ is worked out in
+Appendix \ref{sec:PXYP}.  In total, we find
 \begin{align}
   H_1^{(2)}
-  \EQFS \f1{N\p{N-1}}
-  \sum_{\Delta\ne0} \f{\v v_\Delta\c\v v_\Delta}{\Delta}
-  \sp{{\col{X}\,}^2 - N \col{X^2}},
+  \EQFS \f{N}{N-1} \sum_{X,Y} \sum_{\Delta\ne0}
+  \f{\v v_X^\Delta\c\v v_Y^\Delta}{\Delta}
+  \p{\col{X}\,\col{Y} - \col{XY}},
   \label{eq:H_1_2}
@@
 where $\EQFS$ denotes equality under a restriction to the fully
-symmetric manifold $\M_0$, $\v v_\Delta$ is the projection of $\v v$
-onto the $\Delta$-eigenspace of the matrix $\m h-\diag\v h$, and
-\begin{align}
-  \col{X} \equiv \sum_p X_p,
-  &&
-  \col{X^2} \equiv \sum_p X_p^2.
+symmetric manifold $\M_0$; $\v v_X^\Delta$ is the projection of
+$\v v_X$ onto the $\Delta$-eigenspace of the matrix $\m h-\diag\v h$;
+and for any single-body operator $Z$, we define the spin-averaged
+operator
+\begin{align}
+  \col{Z} \equiv \f1N \sum_p Z_p.
 \end{align}
@@
 \section{Two-body perturbations}
-\label{sec:two_body}
+\label{sec:two_body_pert}
 
@@
 \eqref{eq:perturbations} is simply the restriction of $\V_2$ onto
-$\M_0$ (see Appendix \ref{sec:PXYP}):
-\begin{align}
-  H_2^{(1)}
-  = \P_0 \V_2 \P_0
-  \EQFS \f12 \bar w \p{\col{X}\odot\col{Y} - \col{X\odot Y}},
-\end{align}
-where $A\odot B\equiv\p{AB+BA}/2$ is a symmetrized product of $A$ and
-$B$; and
-\begin{align}
-  \bar w \equiv {N\choose 2}^{-1} \sum_{p<q} w_{pq},
-  &&
-  \col{X} \equiv \sum_p X_p,
-  &&
-  \col{Y} \equiv \sum_p Y_p,
-  &&
-  \col{X\odot Y} \equiv \sum_p X_p\odot Y_p.
+$\M_0$:
+\begin{align}
+  H_2^{(1)} = \sum_O \sum_{p<q} w_{Opq} \P_0 O_{pq} \P_0
+  \EQFS \sum_O \sum_{p<q} w_{Opq} O_{0,1}
+  \EQFS {N \choose 2} \sum_O \bar w_O \col{O},
+\end{align}
+where
+\begin{align}
+  \bar w_O \equiv {N \choose 2}^{-1} \sum_{p<q} w_{Opq},
+  &&
+  \col{O} \equiv {N \choose 2}^{-1} \sum_{p<q} O_{pq},
+\end{align}
+are averages of $w_{Opq}$ and $O_{pq}$ over all pairs of spins $p,q$.
+If the two-body operator $O=X\otimes Y$, then
+\begin{align}
+  \col{O} \EQFS X_0 Y_1
+  \EQFS \f1{N-1} \p{N\, \col{X}\,\col{Y} - \col{XY}},
+\end{align}
+so
+\begin{align}
+  H_2^{(1)} \EQFS \f12 N \sum_{\p{X,Y}} \bar w_{\p{X,Y}}
+  \p{N\, \col{X}\,\col{Y} - \col{XY}}.
 \end{align}
@@
 \begin{align}
-  H_0 \V_2\p{\m w} \ket\psi
-  = E_0 \V_2\p{\m w} \ket\psi
-  + \f12 \sum_{p\ne q} \sp{2\bar h_{pq} w_{pq}
-    + \sum_k \p{h_{pk} w_{kq} + w_{pk} h_{kq}}}
-  X_p Y_q \ket\psi,
+  H_0 \V_2 \ket\psi
+  = E_0 \V_2 \ket\psi
+  + \f12 \sum_{p\ne q} \sp{2\bar h_{pq} w_{Opq}
+    + \sum_k \p{h_{pk} w_{Okq} + w_{Opk} h_{kq}}}
+  O_{pq} \ket\psi,
   \label{eq:H_V2_psi}
@@
   &&
-  h_p \equiv \sum_q h_{pq}.
-\end{align}
-We thus find that the vector $\V_2\p{\m w_\Delta}\ket\psi$ is an
-eigenvector of the interaction Hamiltonian $H_0$ if the coefficients
-$\m w_\Delta$ satisfy the eigenvalue equation
-\begin{align}
-  \check{\m h} \c \v{\m w}_\Delta = \Delta \v{\m w}_\Delta,
+  h_p \equiv \sum_q h_{pq},
+\end{align}
+and we define $w_{Opq}$ for all $p,q$ (as opposed to only $p<q$) by
+enforcing $w_{Opq}=w_{Oqp}$ and $w_{Opp}=0$.  We thus find that the
+vector $\V_2^\Delta\ket\psi$ indexed by a scalar $\Delta$ is an
+eigenvector of the interaction Hamiltonian $H_0$ if its coefficients
+$\v{\m w}_O^\Delta$ all satisfy the eigenvalue equation
+\begin{align}
+  \check{\m h} \c \v{\m w} = \Delta \v{\m w},
   \label{eq:cond_2}
 \end{align}
-where for any symmetric $N\times N$ matrix $\m m$ with zeroes on the
-diagonal (as is the case with the coefficient matrix $\m w_\Delta$),
-we define the ${N \choose 2}$-component vector
-\begin{align}
-  \v{\m m} \equiv \sum_{p<q} m_{pq} \ket{\set{p,q}},
-  \label{eq:flat_vec}
-\end{align}
-and we define the ${N \choose 2} \times {N \choose 2}$ matrix
-\begin{align}
-  \check{\m h} = 2\sum_{p<q} \bar h_{pq} \op{\set{p,q}}
+where
+\begin{align}
+  \check{\m h} \equiv 2\sum_{p<q} \bar h_{pq} \op{\set{p,q}}
   + \sum_{p<q} \ket{\set{p,q}} \sum_{k\notin\set{p,q}}
@@
 \end{align}
-The basis vectors $\ket{\set{p,q}}$ are labeled by a choice
-$\set{p,q}$ of two distinct spins $p,q$.  The eigenvalue problem in
-\eqref{eq:cond_2} can be simplified when the interaction Hamiltonian
-$H_0$ and the perturbation $\V_2\p{\m w}$ are translationally
-invariant; we discuss this simplification in Appendix
-\ref{sec:trans_inv}.
-
-If we {\it construct} a perturbation $\V_2\p{\m w_\Delta}$ with
-coefficients $\m w_\Delta$ that satisfy \eqref{eq:cond_2} with
-eigenvalue $\Delta$, then
-\begin{align}
-  H_0 \V_2\p{\m w_\Delta} \ket\psi
-  = \p{E_0 + \Delta} \V_2\p{\m w_\Delta} \ket\psi,
-\end{align}
-which implies that $\V_2\p{\m w_\Delta}\ket\psi$ is a state of
-definite energy that depends only on the coefficients $\m w_\Delta$.
-As with the case of single-body perturbations, we can decompose any
-matrix $\m w$ into its projections $\m w_\Delta$ onto the eigenspace
-of matrices satisfying \eqref{eq:cond_2} with eigenvalue $\Delta$,
-i.e.~$\m w=\sum_\Delta\m w_\Delta$, which allows us to expand
-\begin{align}
-  \V_2\p{\m w} = \sum_\Delta \V_2\p{\m w_\Delta},
+The eigenvalue problem in \eqref{eq:cond_2} can be simplified somewhat
+when the interaction Hamiltonian $H_0$ is translationally invariant;
+we discuss this simplification in Appendix \ref{sec:trans_inv}.
+
+If we {\it construct} a perturbation $\V_2^\Delta$ with coefficients
+$\v{\m w}_O^\Delta$ that satisfy \eqref{eq:cond_2} with eigenvalue
+$\Delta$, then
+\begin{align}
+  H_0 \V_2^\Delta \ket\psi = \p{E_0 + \Delta} \V_2^\Delta \ket\psi,
+\end{align}
+which implies that $\V_2^\Delta\ket\psi$ is an (unnormalized0 state of
+definite energy that depends only on the coefficients $\m w_O^\Delta$.
+Similarly to the case of single-body perturbations, we can decompose
+any vector $\v{\m w}_O$ into its projections $\v{\m w}_O^\Delta$ onto
+the eigenspace of vectors satisfying \eqref{eq:cond_2} with eigenvalue
+$\Delta$, i.e.~$\v{\m w}_O=\sum_\Delta\v{\m w}_O^\Delta$, which allows
+us to expand
+\begin{align}
+  \V_2 = \sum_\Delta \V_2^\Delta,
+  &&
+  \V_2^\Delta \equiv \sum_O \v{\m w}_O^\Delta \c \v{\m O},
 \end{align}
@@
 \begin{align}
-  H_2^{(2)} = - \P_0 \V_2\p{\m w} \E \V_2\p{\m w} \P_0
-  = -\sum_{\Delta\ne0} \f1\Delta \P_0 \V_2\p{\m w_\Delta}^2 \P_0,
-\end{align}
-where the product $\P_0 \V_2\p{\m w_\Delta}^2 \P_0$ is simplified in
-Appendices \ref{sec:PXYXYP} and \ref{sec:multi_to_collective}.  If
-$X=Y=s_\z$, where $s_\z\equiv\p{\op\up-\op\dn}/2$ is an SU(2) spin
-operator, then
-\begin{align}
-  H_2^{(2)} \EQFS W_4\p{\m h,\m w} S_\z^4 - W_2\p{\m h,\m w} S_\z^2,
+  H_2^{(2)} = - \P_0 \V_2^\Delta \E \V_2^\Delta \P_0
+  = -\sum_{\Delta\ne0} \f1\Delta \P_0 \p{\V_2^\Delta}^2 \P_0,
+\end{align}
+where the product $\P_0 \p{\V_2^\Delta}^2 \P_0$ is simplified in
+Appendices \ref{sec:POQP} and \ref{sec:multi_to_collective}.  In the
+special case that
+\begin{align}
+  \V_2 = \sum_{p<q} w_{pq} s_\z^{(p)} s_\z^{(q)},
+\end{align}
+where $s_\z\equiv\p{\op\up-\op\dn}/2$ is an SU(2) spin-$z$ operator,
+then
+\begin{align}
+  H_2^{(2)} \EQFS W_4 S_\z^4 - W_2 S_\z^2,
 \end{align}
 where $S_\z\equiv\sum_p s_\z^{(p)}$ is a collective SU(2) spin
-operator, $W_k\p{\m h,\m w}$ are scalar coefficients, and we have
-neglected scalar terms in $H_2^{(2)}$ without physical consequence.
-In order to expand the coefficients $W_k\p{\m h,\m w}$, for any
-symmetric matrix $\m m$ with zeros on the diagonal, similarly to the
-special definition $\v{\m m}$ in \eqref{eq:flat_vec} we define
+operator, $W_k$ are scalar coefficients, and we have neglected scalar
+terms in $H_2^{(2)}$ without physical consequence.  In order to expand
+the coefficients $W_k$, for any ${N \choose 2}$-component vector
+$\v{\m m}=\sum_{p<q}m_{pq}\ket{\set{p,q}}$ we define the $N$-component
+vector
 \begin{align}
@@
 \end{align}
-which is essentially the sum over all columns (or rows) of $\m m$, in
-terms of which
-\begin{align}
-  W_4\p{\m h,\m w}
+where we define $m_{pq}$ for all $p,q$ (as opposed to only $p<q$) by
+enforcing $m_{pq}=m_{qp}$ and $m_{pp}=0$.  We then denote the
+projection of $\v{\m m}$ onto the $\Delta$-eigenspace of the matrix
+$\check{\m h}$ defined in \eqref{eq:h_super_mat} by $\v{\m m}_\Delta$,
+which allows us to expand
+\begin{align}
+  W_4
   \equiv \sum_{\Delta\ne0} \f1\Delta \f{\v w_\Delta\c\v w_\Delta
@@
 \end{align}
-\begin{align}
-  W_2\p{\m h,\m w}
+and
+\begin{align}
+  W_2
   \equiv \sum_{\Delta\ne0} \f1{\Delta} \sp{\f{\v w_\Delta\c\v w_\Delta
@@
 \end{align}
-The dependence of the coefficients of $W_k\p{\m h,\m w}$ on the
-interaction matrix $\m h$ is implicit in the decomposition of the
-coefficient matrix $\m w$ into components $\m w_\Delta$ as
-$\m w=\sum_\Delta\m w_\Delta$, where $\m w_\Delta$ is the projection
-of $\m w$ onto the $\Delta$-eigenspace of the matrix $\check{\m h}$
-defined in \eqref{eq:h_super_mat}.
 
@@
   H_0 \V_1 \ket\psi
-  = \sum_{\substack{k\\p<q}} h_{pq} v_k \Pi_{pq} X_k \ket\psi,
+  = \sum_{X,k} \sum_{p<q} h_{pq} v_{Xk} \Pi_{pq} X_k \ket\psi,
   \label{eq:H_V1_psi_start}
@@
 permutation operator $\Pi_{pq}$ commutes with $X_k$ and annihilates on
-$\ket\psi$, leaving us the sum
+$\ket\psi$, leaving terms at each fixed $k$ of the form
 \begin{align}
   \sum_{\substack{p<q\\k\notin\set{p,q}}} h_{pq}
-  = \f12 \sum_{\substack{p,q\\k\notin\set{p,q}}} h_{pq}
-  = \f12 \sum_{p,q} h_{pq}
-  - \f12 \sum_{\substack{p,q\\k\in\set{p,q}}} h_{pq}
+  = \sum_{p<q} h_{pq} - \sum_{\substack{p<q\\k\in\set{p,q}}} h_{pq}
+  = E_0 - \f12 \sum_{\substack{p,q\\k\in\set{p,q}}} h_{pq}
   = E_0 - h_k,
 \end{align}
-where we have used the facts that $h_{pq}=h_{qp}$ and $h_{pp}=0$.  The
-terms in \eqref{eq:H_V1_psi_start} with $k\notin\set{p,q}$ are then
-\begin{align}
-  \sum_{\substack{p<q\\k\notin\set{p,q}}}
-  h_{pq} v_k \Pi_{pq} X_k \ket\psi
-  = \sum_k \p{E_0 - h_k} v_k X_k \ket\psi
-  = E_0 \V_1 \ket\psi - \sum_k h_k v_k X_k \ket\psi
-\end{align}
-while the terms with $k\in\set{p,q}$ are
+where we define $h_{pq}$ for all $p,q$ (as opposed to only $p<q$) by
+enforcing $h_{pq}=h_{qp}$ and $h_{pp}=0$; and
+$h_k \equiv \sum_\ell h_{k\ell}$.  The terms in
+\eqref{eq:H_V1_psi_start} with $k\notin\set{p,q}$ are then
+\begin{align}
+  \sum_X \sum_{\substack{p<q\\k\notin\set{p,q}}}
+  h_{pq} v_{Xk} \Pi_{pq} X_k \ket\psi
+  = \sum_{X,k} \p{E_0 - h_k} v_{Xk} X_k \ket\psi
+  = E_0 \V_1 \ket\psi - \sum_{X,k} h_k v_{Xk} X_k \ket\psi
+\end{align}
+while the terms in \eqref{eq:H_V1_psi_start} with $k\in\set{p,q}$ for
+each fixed $X$ take the form
 \begin{align}
   \sum_{\substack{p<q\\k\in\set{p,q}}}
-  h_{pq} v_k \Pi_{pq} X_k \ket\psi
-  = \sum_{k<q} h_{kq} v_k X_q \ket\psi
-  + \sum_{p<k} h_{pk} v_k X_p \ket\psi
-  = \sum_{k,q} h_{kq} v_k X_q \ket\psi
+  h_{pq} v_{Xk} \Pi_{pq} X_k \ket\psi
+  = \sum_{k<q} h_{kq} v_{Xk} X_q \ket\psi
+  + \sum_{p<k} h_{pk} v_{Xk} X_p \ket\psi
+  = \sum_{k,q} h_{kq} v_{Xk} X_q \ket\psi
 \end{align}
@@
   = E_0 \V_1 \ket\psi
-  + \p{\sum_{p,q} h_{pq} v_p X_q - \sum_q h_q v_q X_q} \ket\psi.
+  + \sum_X \p{\sum_{p,q} h_{pq} v_{Xp} X_q - \sum_p h_p v_{Xp} X_p}
+  \ket\psi.
 \end{align}
@@
 \begin{align}
-  H_0 \V_2\p{\m w} \ket\psi
-  = \f12 \sum_{\substack{k,\ell\\p<q}} w_{k\ell} h_{pq}
-  \Pi_{pq} X_k Y_\ell \ket\psi,
+  H_0 \V_2 \ket\psi
+  = \sum_O \sum_{\substack{k<\ell\\p<q}} w_{Ok\ell} h_{pq}
+  \Pi_{pq} O_{k\ell} \ket\psi,
   \label{eq:H_V2_psi_start}
 \end{align}
-which has one term with $\set{p,q}=\set{k,\ell}$, terms with
-$p,q\notin\set{k,\ell}$, and terms with only one of
-$p,q\in\set{k,\ell}$.  The permutation operator $\Pi_{pq}$ acts
-trivially on $X_k Y_\ell \ket\psi$ when $\set{p,q}=\set{k,\ell}$ or
-$p,q\notin\set{k,\ell}$, leaving a sum over $p,q$ of the form
+which has terms with $\set{p,q}=\set{k,\ell}$, terms with
+$p,q\notin\set{k,\ell}$, and terms with only one of $p$ or
+$q\in\set{k,\ell}$.  The permutation operator $\Pi_{pq}$ acts
+trivially on $O_{k_\ell}\ket\psi$ when $\set{p,q}=\set{k,\ell}$ or
+$p,q\notin\set{k,\ell}$, leaving terms at each fixed $k,\ell$ of the
+form
 \begin{align}
@@
   - \sum_{\substack{p\in\set{k,\ell}\\q\notin\set{k,\ell}}} h_{pq}
-  = E_0 + 2 h_{k\ell} - h_k - h_\ell
+  = E_0 - h_k - h_\ell + 2 h_{k\ell}
   = E_0 + 2 \bar h_{k\ell},
@@
 The corresponding terms in \eqref{eq:H_V2_psi_start} with
-$p,q\notin\set{k,\ell}$ and $\set{p,q}=\set{k,\ell}$ are then
+$\set{p,q}=\set{k,\ell}$ or $p,q\notin\set{k,\ell}$ are then
 \begin{align}
   E_0 \V_2 \ket\psi
-  + \sum_{k,\ell} \bar h_{k\ell} w_{k\ell} X_k Y_\ell \ket\psi.
+  + \sum_O \sum_{k<\ell} 2 \bar h_{k\ell} w_{Ok\ell} O_{k\ell} \ket\psi.
 \end{align}
@@
 \begin{align}
-  \f12 \sum_{k\ne\ell} \sum_q \p{h_{kq} w_{q\ell} + w_{kq} h_{q\ell}}
-  X_k Y_\ell \ket\psi.
+  \f12 \sum_{k\ne\ell} \sum_q \p{h_{kq} w_{Oq\ell} + w_{Okq} h_{q\ell}}
+  O_{k\ell} \ket\psi.
 \end{align}
@@
   &= E_0 \V_2 \ket\psi
-  + \sum_{k,\ell} \bar h_{k\ell} w_{k\ell} X_k Y_\ell \ket\psi
-  + \f12 \sum_{k\ne\ell} \sum_q \p{h_{kq} w_{q\ell} + w_{kq} h_{q\ell}}
-  X_k Y_\ell \ket\psi \\
+  + \sum_{k<\ell} 2\bar h_{k\ell} w_{Ok\ell} O_{k\ell} \ket\psi
+  + \f12 \sum_{k\ne\ell} \sum_q \p{h_{kq} w_{Oq\ell} + w_{Okq} h_{q\ell}}
+  O_{k\ell} \ket\psi \\
   &= E_0 \V_2 \ket\psi
-  + \f12 \sum_{k\ne\ell} \sp{2\bar h_{k\ell} w_{k\ell}
-    + \sum_q \p{h_{kq} w_{q\ell} + w_{kq} h_{q\ell}}}
-  X_k Y_\ell \ket\psi
+  + \f12 \sum_{k\ne\ell} \sp{2\bar h_{k\ell} w_{Ok\ell}
+    + \sum_q \p{h_{kq} w_{Oq\ell} + w_{Okq} h_{q\ell}}}
+  O_{k\ell} \ket\psi.
 \end{align}
@@
 \subsection{Single-body product}
-\label{sec:PXXP}
+\label{sec:PXYP}
 
@@
 \begin{align}
-  \P_0 \V_1\p{\v v}^2 \P_0
-  \EQFS \sum_{p,q} v_p v_q X_p X_q
-  = \sum_p v_p^2 X_p^2
-  + \sum_{p\ne q} v_p v_q X_p X_q,
-  \label{eq:PXXP_start}
+  \P_0 \p{\V_1^\Delta}^2 \P_0
+  \EQFS \sum_{\substack{X,Y\\p,q}} v_{Xp} v_{Yq} X_p Y_q
+  = \sum_{X,Y,p} v_{Xp} v_{Yp} X_p Y_p
+  + \sum_{\substack{X,Y\\p\ne q}} v_{Xp} v_{Yq} X_p Y_q,
+  \label{eq:PXYP_start}
 \end{align}
 where $\EQFS$ denotes equality under a restriction to the fully
-symmetric manifold $\M_0$; and the coefficients $\v v$ satisfy
-\eqref{eq:cond_1} with eigenvalue $\Delta\ne0$, which implies that
-$\v v$ is mean-zero.  The first sum on the right of
-\eqref{eq:PXXP_start} is then
-\begin{align}
-  \sum_p v_p^2 X_p^2
-  \EQFS \sum_p v_p^2 X_0^2
-  \EQFS \f1N \v v\c\v v\, \col{X^2},
-  &&
-  \col{X^2} \equiv \sum_p X_p^2.
-  \label{eq:PXXP_eq}
-\end{align}
-The second sum on the right of \eqref{eq:PXXP_start} is
-\begin{align}
-  \sum_{p\ne q} v_p v_q X_p X_q
-  \EQFS \sum_{p\ne q} v_p v_q X_0 X_1,
-  \label{eq:PXXP_neq_start}
-\end{align}
-where the sum over coefficients in \eqref{eq:PXXP_neq_start} can be
-simplified to
-\begin{align}
-  \sum_{p\ne q} v_p v_q
-  = \sum_{p,q} v_p v_q - \sum_{p=q} v_p v_q
-  = - \v v\c\v v
+symmetric manifold $\M_0$; and all coefficient vectors $\v v_X$
+satisfy \eqref{eq:cond_1} with eigenvalue $\Delta\ne0$, which implies
+that all $\v v_X$ are mean-zero.  The first sum on the right of
+\eqref{eq:PXYP_start} is then
+\begin{align}
+  \sum_{X,Y,p} v_{Xp} v_{Yp} X_p Y_p
+  \EQFS \sum_{X,Y,p} v_{Xp} v_{Yp} X_0 Y_0
+  \EQFS \sum_{X,Y} \v v_X \c\v v_Y \col{XY},
+  &&
+  \col{Z} \equiv \f1N \sum_p Z_p.
+  \label{eq:PXYP_eq}
+\end{align}
+The second sum on the right of \eqref{eq:PXYP_start}, meanwhile, is
+\begin{align}
+  \sum_{\substack{X,Y\\p\ne q}} v_{Xp} v_{Yq} X_p Y_q
+  \EQFS \sum_{\substack{X,Y\\p\ne q}} v_{Xp} v_{Yq} X_0 Y_1,
+  \label{eq:PXYP_neq_start}
+\end{align}
+where we can use the fact that the vectors $\v v_X,\v v_Y$ are
+mean-zero to simplify the sum over $p\ne q$ in
+\eqref{eq:PXYP_neq_start} to
+\begin{align}
+  \sum_{p\ne q} v_{Xp} v_{Yq}
+  = \sum_{p,q} v_{Xp} v_{Yq} - \sum_{p=q} v_{Xp} v_{Yq}
+  = - \v v_X\c\v v_Y,
 \end{align}
 In order to simplify the product of operators in
-\eqref{eq:PXXP_neq_start}, we expand
-\begin{align}
-  {\col{X}\,}^2
-  = \sum_p X_p^2 + \sum_{p\ne q} X_p X_q
-  \EQFS \col{X^2} + N \p{N-1} X_0 X_1,
-  \label{eq:PXXP_neq_ops}
+\eqref{eq:PXYP_neq_start}, we expand
+\begin{align}
+  N^2 \col{X}\,\col{Y} = \sum_{p,q} X_p Y_q
+  = \sum_{p=q} X_p Y_q + \sum_{p\ne q} X_p Y_q
+  \EQFS N \col{XY} + N\p{N-1} X_0 Y_1,
+  \label{eq:PXYP_neq_ops}
 \end{align}
@@
 \begin{align}
-  \sum_{p\ne q} v_p v_q X_p X_q
-  \EQFS - \v v\c\v v X_0 X_1
-  \EQFS -\f{\v v\c\v v}{N\p{N-1}}
-  \sp{{\col{X}\,}^2 - \col{X^2}}
-  \label{eq:PXXP_neq}
+  \sum_{\substack{X,Y\\p\ne q}} v_{Xp} v_{Yq} X_p Y_q
+  \EQFS - \sum_{X\ne Y} \v v_X\c\v v_Y X_0 Y_1
+  \EQFS - \f1{N-1}\sum_{X,Y} \v v_X\c\v v_Y
+  \p{N \col{X}\,\col{Y} - \col{XY}}.
+  \label{eq:PXYP_neq}
 \end{align}
@@
 \begin{align}
-  \P_0 \V_1\p{\v v}^2 \P_0
-  \EQFS -\f{\v v\c\v v}{N\p{N-1}}
-  \sp{{\col{X}\,}^2 - N \col{X^2}}.
+  \P_0 \p{\V_1^\Delta}^2 \P_0
+  = - \f{N}{N-1} \sum_{X,Y} \v v_X \c\v v_Y
+  \p{\col{X}\,\col{Y} - \col{XY}}.
 \end{align}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Two-body product, first order}
-\label{sec:PXYP}
-
-Calculating the first-order effective Hamiltonian $H_2^{(1)}$ induced
-on the fully symmetric manifold $\M_0$ by the perturbation $\V_2$ in
-\eqref{eq:perturbations} requires us to simplify
-\begin{align}
-  \P_0 \V_2 \P_0
-  \EQFS \f12 \sum_{p\ne q} w_{pq} X_p Y_q
-  = {N\choose 2} \bar w X_0 Y_1,
-  \label{eq:PXYP_start}
-\end{align}
-where $\EQFS$ denotes equality under a restriction to the fully
-symmetric manifold $\M_0$, ${N\choose2}=N\p{N-1}/2$ is a binomial
-coefficient, and
-\begin{align}
-  \bar w \equiv {N\choose 2}^{-1} \sum_{p<q} w_{pq}
-\end{align}
-is the mean value of the pair-wise couplings in $\V_2$.  To simplify
-the operator content of \eqref{eq:PXYP_start}, we define
-\begin{align}
-  \col{X} \equiv \sum_p X_p,
-  &&
-  \col{Y} \equiv \sum_p Y_p,
-  &&
-  A \odot B \equiv \f12\p{AB+BA},
-  &&
-  \col{X\odot Y} \equiv \sum_p X_p \odot Y_p,
-\end{align}
-and expand
-\begin{align}
-  \col{X}\odot\col{Y} \EQFS \col{X\odot Y} + N\p{N-1} X_0 Y_1.
-\end{align}
-Substituting the product $X_0 Y_1$ back into \eqref{eq:PXYP_start}
-gives us
-\begin{align}
-  \P_0 \V_2\p{\m w} \P_0
-  \EQFS \f12 \bar w \p{\col{X}\odot\col{Y} - \col{X\odot Y}}.
-\end{align}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Two-body product, second order}
-\label{sec:PXYXYP}
+\subsection{Two-body product}
+\label{sec:POQP}
 
@@
 \begin{align}
-  \P_0 \V_2\p{\m w}^2 \P_0
-  \EQFS \f14 \sum_{k,\ell,p,q}
-  w_{k\ell} w_{pq} X_k Y_\ell X_p Y_q,
-  \label{eq:PXYXYP_start}
+  \P_0 \p{\V_2^\Delta}^2 \P_0
+  = \sum_{O,Q} \sum_{\substack{k<\ell\\p<q}}
+  w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq}
+  \EQFS \f14 \sum_{O,Q} \sum_{k,\ell,p,q}
+  w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq},
+  \label{eq:POQP_start}
 \end{align}
 where $\EQFS$ denotes equality under a restriction to the fully
-symmetric manifold $\M_0$, and the coefficients $\m w$ satisfy
-\eqref{eq:cond_2} with eigenvalue $\Delta\p{\m w}\ne0$, which implies
-that $\m w$ is mean-zero.  To simplify the sum in
-\eqref{eq:PXYXYP_start}, we first consider the terms with
-$p,q\in\set{k,\ell}$:
-\begin{align}
-  \sum_{\substack{k,\ell\\p,q\in\set{k,\ell}}}
-  w_{k\ell} w_{pq} X_k Y_\ell X_p Y_q
-  \EQFS 2 \v{\m w}\c\v{\m w} \p{X_0^2 Y_1^2 + X_0 Y_0 Y_1 X_1},
-  &&
-  \v{\m w} \equiv \sum_{k<\ell} w_{k\ell} \ket{\set{k,\ell}}.
-\end{align}
-The terms of the sum in \eqref{eq:PXYXYP_start} with
-$p\in\set{k,\ell}$ and $q\notin\set{k,\ell}$ or vice versa are
-\begin{align}
-  \sum_{\substack{k,\ell\\p\in\set{k,\ell}\\q\notin\set{k,\ell}}}
-  w_{k\ell} w_{pq} X_k Y_\ell X_p Y_q
-  &\EQFS \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}}
-  w_{k\ell} w_{k q} \p{X_0^2 Y_1 Y_2 + Y_0 X_0 X_1 Y_2}, \\
-  \sum_{\substack{k,\ell\\p\notin\set{k,\ell}\\q\in\set{k,\ell}}}
-  w_{k\ell} w_{pq} X_k Y_\ell X_p Y_q
-  &\EQFS \sum_{\substack{k,\ell\\p\notin\set{k,\ell}}}
-  w_{k\ell} w_{pk} \p{X_0 Y_0 X_1 Y_2 + Y_0^2 X_1 X_2},
+symmetric manifold $\M_0$; we define $w_{Opq}$ for all $p,q$ (as
+opposed to only $p<q$) by enforcing $w_{Opq}=w_{Oqp}$ and $w_{Opp}=0$;
+and the coefficient vectors $\v w_O$ satisfy the eigenvalue equation
+\eqref{eq:cond_2} with eigenvalue $\Delta\ne0$, which implies that all
+$\v{\m w}_O$ are mean-zero.  To simplify the sum in
+\eqref{eq:POQP_start}, we first consider the terms with
+$\p{p,q}=\p{k,\ell}$:
+\begin{align}
+  \sum_{k<\ell}
+  w_{Ok\ell} w_{Qk\ell} O_{k\ell} Q_{k\ell}
+  = \v{\m w}_O \c \v{\m w}_Q O_{0,1} Q_{0,1}.
+\end{align}
+We then consider terms $p\in\set{k,\ell}$ and $q\notin\set{k,\ell}$ or
+vice versa:
+\begin{align}
+  \f14 \sum_{\substack{k,\ell\\p\in\set{k,\ell}\\q\notin\set{k,\ell}}}
+  w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq}
+  &\EQFS \f14 \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}}
+  w_{Ok\ell} w_{Qkq} \p{O_{0,1} Q_{0,2} + O_{0,1} Q_{1,2}} \\
+  \f14 \sum_{\substack{k,\ell\\p\notin\set{k,\ell}\\q\in\set{k,\ell}}}
+  w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq}
+  &\EQFS \f14 \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}}
+  w_{Ok\ell} w_{Qkq} \p{O_{0,1} Q_{2,0} + O_{0,1} Q_{2,1}},
 \end{align}
@@
 \begin{align}
-  \sum_{\substack{k,\ell\\p\notin\set{k,\ell}}} w_{k\ell} w_{pk}
-  = \sum_{k,\ell,p} w_{k\ell} w_{kp} - \sum_{k,\ell} w_{k\ell}^2
-  = \v w\c\v w - 2 \v{\m w}\c\v{\m w},
-  &&
-  \v w \equiv \sum_{k,\ell} w_{k\ell} \ket{k}.
-\end{align}
-Finally, the terms of the sum in \eqref{eq:PXYXYP_start} with
+  \f14 \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}} w_{Ok\ell} w_{Qkq}
+  = \f14 \sum_{k,\ell,q} w_{Ok\ell} w_{Qkq}
+  - \f14 \sum_{k,\ell} w_{Ok\ell} w_{Qk\ell}
+  = \f14 \v w_O\c\v w_Q - \f12 \v{\m w}_O \c \v{\m w}_Q.
+\end{align}
+Finally, the terms in \eqref{eq:POQP_start} with
 $p,q\notin\set{k,\ell}$ are
 \begin{align}
-  \sum_{\substack{k,\ell\\p,q\notin\set{k,\ell}}}
-  w_{k\ell} w_{pq} X_k Y_\ell X_p Y_q
-  \EQFS \sum_{\substack{k,\ell\\p,q\notin\set{k,\ell}}}
-  w_{k\ell} w_{pq} X_0 X_1 Y_2 Y_3,
-\end{align}
-where, using the fact that $\m w$ is mean-zero,
-\begin{align}
-  \sum_{\substack{k,\ell\\p,q\notin\set{k,\ell}}} w_{k\ell} w_{pq}
-  = - \sum_{\substack{k,\ell\\p,q\in\set{k,\ell}}} w_{k\ell} w_{pq}
-  - \sum_{\substack{k,\ell\\p\in\set{k,\ell}\\q\notin\set{k,\ell}}}
-  w_{k\ell} w_{pq}
-  - \sum_{\substack{k,\ell\\p\notin\set{k,\ell}\\q\in\set{k,\ell}}}
-  w_{k\ell} w_{pq}
-  = -4 \v w\c\v w + 4 \v{\m w}\c\v{\m w}.
-\end{align}
+  \f14 \sum_{\substack{k,\ell\\p,q\notin\set{k,\ell}}}
+  w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq}
+  \EQFS \f14 \sum_{\substack{k,\ell\\p,q\notin\set{k,\ell}}}
+  w_{Ok\ell} w_{Qpq} O_{0,1} Q_{2,3},
+\end{align}
+where, using the fact that all $\v{\m w}_O$ are mean-zero,
+\begin{multline}
+  \f14 \sum_{\substack{k,\ell\\p,q\notin\set{k,\ell}}} w_{Ok\ell} w_{Qpq}
+  = - \f14 \sum_{\substack{k,\ell\\p,q\in\set{k,\ell}}} w_{Ok\ell} w_{Qpq}
+  - \f14 \sum_{\substack{k,\ell\\p\in\set{k,\ell}\\q\notin\set{k,\ell}}}
+  w_{Ok\ell} w_{Qpq}
+  - \f14 \sum_{\substack{k,\ell\\p\notin\set{k,\ell}\\q\in\set{k,\ell}}}
+  w_{Ok\ell} w_{Qpq} \\
+  = -\v w_O \c \v w_Q + \v{\m w}_O \c \v{\m w}_Q.
+\end{multline}
 Altogether, we thus have that
 \begin{multline}
-  \P_0 \V_2\p{\m w}^2 \P_0
-  \EQFS -\p{\v w\c\v w - \v{\m w}\c\v{\m w}} X_0 X_1 Y_2 Y_3
-  + \f12 \v{\m w}\c\v{\m w} \p{X_0^2 Y_1^2 + X_0 Y_0 Y_1 X_1} \\
-  + \f14 \p{\v w\c\v w - 2\v{\m w}\c\v{\m w}}
-  \p{X_0^2 Y_1 Y_2 + X_0 Y_0 X_1 Y_2
-    + Y_0 X_0 X_1 Y_2 + Y_0^2 X_1 X_2}.
-  \label{eq:PXYXYP}
+  \P_0 \p{\V_2^\Delta}^2 \P_0
+  \EQFS - \sum_{O,Q} \bigg[
+  \p{\v w_O \c \v w_Q - \v{\m w}_O \c \v{\m w}_Q} O_{0,1} Q_{2,3}
+  - \v{\m w}_O \c \v{\m w}_Q O_{0,1} Q_{0,1} \\
+  -\f14 \p{\v w_O\c\v w_Q - 2 \v{\m w}_O \c \v{\m w}_Q}
+  O_{0,1} \p{Q_{0,2} + Q_{2,0} + Q_{1,2} + Q_{2,1}} \bigg].
+  \label{eq:POQP}
 \end{multline}
@@
 turns out to be rather non-trivial in general.  We therefore discuss
-the decomposition of multi-local operators (e.g.~$X_0Y_0X_1Y_2$, a
+the decomposition of multi-local operators (e.g.~$O_{0,1} Q_{1,2}$, a
 3-local operator) into a sum of collective operators within the fully
@@
 Here we simplify the eigenvalue problems that appear at second order
-in the single- and two-body theory in Sections \ref{sec:single_body}
-and \ref{sec:two_body} in the case of a translationally invariant
-system with periodic boundary conditions.  We solve the single-body
-eigenvalue problem analytically, and reduce the two-body eigenvalue
-problem to that of diagonalizing a single $\p{N-1}\times\p{N-1}$
-matrix.
+in the single- and two-body theory in Sections
+\ref{sec:single_body_pert} and \ref{sec:two_body_pert} in the case of
+a translationally invariant system with periodic boundary conditions.
+We solve the single-body eigenvalue problem analytically, and reduce
+the two-body eigenvalue problem to that of diagonalizing an
+$\floor{N/2}\times\floor{N/2}$ matrix.
 
@@
 where we place a bar over a single-spin operator to denote its
-collective version, i.e.~$\col{Q}\equiv\sum_j Q^{(j)}$, where
-$Q^{(j)}$ is an operator that acts with the single-spin operator $Q$
-on spin $j$ and trivially on all other spins.  By construction, the
-set of collective operators $\set{\O_P:\t{partitions}~P~\t{of}~S}$ is
-one-to-one to the set of multi-local operators
-$\set{O_P:\t{partitions}~P~\t{of}~S}$, and it is straightforward to
-work out that for any set $S$ of single-spin operators,
-\begin{align}
-  \O_S
+spin-average, i.e.~$\col{Q}\equiv\sum_j Q^{(j)}/N$, where $Q^{(j)}$
+acts with the single-body operator $Q$ on spin $j$ and trivially on
+all other spins.  By construction, the set of collective operators
+$\set{\O_P:\t{partitions}~P~\t{of}~S}$ is one-to-one to the set of
+multi-local operators $\set{O_P:\t{partitions}~P~\t{of}~S}$, and it is
+straightforward to work out that for any set $S$ of single-spin
+operators,
+\begin{align}
+  N^{\abs{S}} \O_S
   \EQFS \sum_{\substack{\t{partitions}\\P~\t{of}~S}}
@@
 \begin{align}
-  N!_{\abs{S}}\, O_S
+  N!_{\abs{S}} O_S
   &\EQFS \sum_{\substack{\t{partitions}\\P~\t{of}~S}}
-  \p{-1}^{\abs{S}-\abs{P}} \O_P \times \prod_{p\in P}\p{\abs{p}-1}!
+  \p{-1}^{\abs{S}-\abs{P}} N^{\abs{P}} \O_P
+  \times \prod_{p\in P}\p{\abs{p}-1}!
   \label{eq:multi_to_collective_full} \\
@@
   \sum_{\substack{\t{ordered}\\\t{partitions}\\\tilde P~\t{of}~\Pi}}
-  \p{-1}^{\abs{S}-\abs{\tilde P}} \O_{\tilde P}
+  \p{-1}^{\abs{S}-\abs{\tilde P}} N^{\abs{\tilde P}} \O_{\tilde P}
   \times \prod_{\tilde p\in\tilde P} \p{\abs{\tilde p}-1}!
@@
 \begin{align}
-  \O_{\tilde S} \EQFS
+  N^{\abs{\tilde S}} \O_{\tilde S} \EQFS
   \sum_{\substack{\t{ordered}\\\t{partitions}\\\tilde P~\t{of}~\tilde S}}
@@
 \eqref{eq:multi_to_collective_full_simp} closely resemble each other,
-in practice the latter involves less computational overhead due to the
-lack of symmetrization in the construction of $\O_{\tilde P}$.  There
-are, in principle, even more economical decompositions of a
-multi-local operators into collective opreators, for example
+the latter can involve less computational overhead due to the lack of
+symmetrization in the construction of $\O_{\tilde P}$.  There are, in
+principle, even more economical decompositions of a multi-local
+operators into collective opreators, for example
 \begin{align}
   N!_2\, X\otimes Y
-  &\EQFS \col{X}\,\col{Y} - \col{XY}, \\
+  &\EQFS N^2\, \col{X}\,\col{Y} - N\, \col{XY}, \\
   N!_3\, X\otimes Y \otimes Z
-  &\EQFS \col{X}\,\col{Y}\,\col{Z}
-  - \col{X}\,\col{YZ} - \col{Y}\,\col{XZ} - \col{Z}\,\col{XY}
-  + \col{ZXY} + \col{YXZ},
+  &\EQFS N^3\, \col{X}\,\col{Y}\,\col{Z}
+  - N^2 \p{\col{X}\,\col{YZ} + \col{Y}\,\col{XZ} + \col{Z}\,\col{XY}}
+  + N \p{\col{ZXY} + \col{YXZ}},
 \end{align}
@@
 \begin{align}
-  N!_2\, X\otimes X \EQFS {\col{X}\,}^2 - \col{X^2},
-  &&
-  N!_3\, X^{\otimes 3} \EQFS {\col{X}\,}^3
-  - 3 \col{X} \odot \col{X^2} + 2\col{X^3},
+  N!_2\, X\otimes X \EQFS N^2\, {\col{X}\,}^2 - N\, \col{X^2},
+  &&
+  N!_3\, X^{\otimes 3} \EQFS N^3\, {\col{X}\,}^3
+  - 3 N^2\, \col{X} \odot \col{X^2} + 2 N\, \col{X^3},
 \end{align}
@@
   N!_4\, X^{\otimes 4}
-  \EQFS {\col{X}\,}^4 - 4 {\col{X}\,}^2 \odot \col{X^2}
-  - 2 \col{X} \, {\col{X}\,}^2 \, \col{X}
-  + 3 {\col{X^2}\,}^2 + 8 \col{X} \odot \col{X^3}
-  - 6 \col{X^4}.
+  \EQFS N^4\, {\col{X}\,}^4 - N^3 \p{4 {\col{X}\,}^2 \odot \col{X^2}
+    + 2 \col{X} \, {\col{X}\,}^2 \, \col{X}}
+  + N^3\p{3 {\col{X^2}\,}^2 + 8 \col{X} \odot \col{X^3}}
+  - 6 N\, \col{X^4}.
 \end{align}
@@
 \begin{align}
-  N!_2\, X\otimes X \EQFS {\col{X}\,}^2 - N,
-  &&
-  N!_3\, X^{\otimes 3} \EQFS {\col{X}\,}^3 - \p{3N-2} \col{X},
-\end{align}
-\begin{align}
-  N!_4\, X^{\otimes 4} \EQFS {\col{X}\,}^4
-  - 2\p{3N-4} {\col{X}\,}^2 + 3 N \p{N-2}.
+  N!_2\, X\otimes X \EQFS N^2\, {\col{X}\,}^2 - N,
+  &&
+  N!_3\, X^{\otimes 3}
+  \EQFS N^3\, {\col{X}\,}^3 - \p{3N-2} N\, \col{X},
+\end{align}
+\begin{align}
+  N!_4\, X^{\otimes 4} \EQFS N^4\, {\col{X}\,}^4
+  - 2\p{3N-4} N^2\, {\col{X}\,}^2 + 3 N \p{N-2}.
 \end{align}

### 2020-01-13 collective_sun/sun_perturbation.tex
revert to previous convention for denoting collective operators, making them extensive

@@
   \EQFS \sum_{X,p} v_{Xp} X_0
-  \EQFS N \sum_X \bar v_X \col{X},
+  \EQFS \sum_X \bar v_X \col{X},
   \label{eq:H_1_1}
@@
   &&
-  \col{X} \equiv \f1N \sum_p X_p.
-\end{align}
-In order to compute the second order effective Hamiltonian
+  \col{X} \equiv \sum_p X_p,
+\end{align}
+are respectively the mean value in $\v v_X$ and the collective version
+of $X$.  In order to compute the second order effective Hamiltonian
 $H_1^{(2)}$, we first choose an arbitrary fully symmetric state
@@
   H_1^{(2)}
-  \EQFS \f{N}{N-1} \sum_{X,Y} \sum_{\Delta\ne0}
+  \EQFS \f1{N\p{N-1}} \sum_{X,Y} \sum_{\Delta\ne0}
   \f{\v v_X^\Delta\c\v v_Y^\Delta}{\Delta}
-  \p{\col{X}\,\col{Y} - \col{XY}},
+  \p{\col{X}\,\col{Y} - N\, \col{XY}},
   \label{eq:H_1_2}
@@
 $\v v_X$ onto the $\Delta$-eigenspace of the matrix $\m h-\diag\v h$;
-and for any single-body operator $Z$, we define the spin-averaged
-operator
-\begin{align}
-  \col{Z} \equiv \f1N \sum_p Z_p.
+and for any single-body operator $Z$, we define the collective operator
+\begin{align}
+  \col{Z} \equiv \sum_p Z_p.
 \end{align}
@@
   \EQFS \sum_O \sum_{p<q} w_{Opq} O_{0,1}
-  \EQFS {N \choose 2} \sum_O \bar w_O \col{O},
+  \EQFS \sum_O \bar w_O \col{O},
 \end{align}
@@
   &&
-  \col{O} \equiv {N \choose 2}^{-1} \sum_{p<q} O_{pq},
-\end{align}
-are averages of $w_{Opq}$ and $O_{pq}$ over all pairs of spins $p,q$.
-If the two-body operator $O=X\otimes Y$, then
-\begin{align}
-  \col{O} \EQFS X_0 Y_1
-  \EQFS \f1{N-1} \p{N\, \col{X}\,\col{Y} - \col{XY}},
-\end{align}
-so
-\begin{align}
-  H_2^{(1)} \EQFS \f12 N \sum_{\p{X,Y}} \bar w_{\p{X,Y}}
-  \p{N\, \col{X}\,\col{Y} - \col{XY}}.
+  \col{O} \equiv \sum_{p<q} O_{pq},
+\end{align}
+are an average of $w_{Opq}$ and a collective version of the two-body
+operator $O_{pq}$.  If the two-body operator $O=X\otimes Y$, then
+\begin{align}
+  H_2^{(1)} \EQFS \f12 \sum_{\p{X,Y}} \bar w_{\p{X,Y}}
+  \p{\col{X}\,\col{Y} - \col{XY}}.
 \end{align}
@@
 \end{align}
-which implies that $\V_2^\Delta\ket\psi$ is an (unnormalized0 state of
+which implies that $\V_2^\Delta\ket\psi$ is an (unnormalized state of
 definite energy that depends only on the coefficients $\m w_O^\Delta$.
@@
   \EQFS \sum_{X,Y,p} v_{Xp} v_{Yp} X_0 Y_0
-  \EQFS \sum_{X,Y} \v v_X \c\v v_Y \col{XY},
-  &&
-  \col{Z} \equiv \f1N \sum_p Z_p.
+  \EQFS \f1N \sum_{X,Y} \v v_X \c\v v_Y \col{XY},
+  &&
+  \col{Z} \equiv \sum_p Z_p.
   \label{eq:PXYP_eq}
@@
 \begin{align}
-  N^2 \col{X}\,\col{Y} = \sum_{p,q} X_p Y_q
+  \col{X}\,\col{Y} = \sum_{p,q} X_p Y_q
   = \sum_{p=q} X_p Y_q + \sum_{p\ne q} X_p Y_q
-  \EQFS N \col{XY} + N\p{N-1} X_0 Y_1,
+  \EQFS \col{XY} + N\p{N-1} X_0 Y_1,
   \label{eq:PXYP_neq_ops}
@@
   \EQFS - \sum_{X\ne Y} \v v_X\c\v v_Y X_0 Y_1
-  \EQFS - \f1{N-1}\sum_{X,Y} \v v_X\c\v v_Y
-  \p{N \col{X}\,\col{Y} - \col{XY}}.
+  \EQFS - \f1{N\p{N-1}}\sum_{X,Y} \v v_X\c\v v_Y
+  \p{\col{X}\,\col{Y} - \col{XY}}.
   \label{eq:PXYP_neq}
@@
   \P_0 \p{\V_1^\Delta}^2 \P_0
-  = - \f{N}{N-1} \sum_{X,Y} \v v_X \c\v v_Y
-  \p{\col{X}\,\col{Y} - \col{XY}}.
+  = - \f1{N\p{N-1}} \sum_{X,Y} \v v_X \c\v v_Y
+  \p{\col{X}\,\col{Y} - N\,\col{XY}}.
 \end{align}
@@
 where we place a bar over a single-spin operator to denote its
-spin-average, i.e.~$\col{Q}\equiv\sum_j Q^{(j)}/N$, where $Q^{(j)}$
-acts with the single-body operator $Q$ on spin $j$ and trivially on
-all other spins.  By construction, the set of collective operators
-$\set{\O_P:\t{partitions}~P~\t{of}~S}$ is one-to-one to the set of
-multi-local operators $\set{O_P:\t{partitions}~P~\t{of}~S}$, and it is
-straightforward to work out that for any set $S$ of single-spin
-operators,
-\begin{align}
-  N^{\abs{S}} \O_S
+collective version, i.e.~$\col{Q}\equiv\sum_j Q^{(j)}$, where
+$Q^{(j)}$ is an operator that acts with the single-spin operator $Q$
+on spin $j$ and trivially on all other spins.  By construction, the
+set of collective operators $\set{\O_P:\t{partitions}~P~\t{of}~S}$ is
+one-to-one to the set of multi-local operators
+$\set{O_P:\t{partitions}~P~\t{of}~S}$, and it is straightforward to
+work out that for any set $S$ of single-spin operators,
+\begin{align}
+  \O_S
   \EQFS \sum_{\substack{\t{partitions}\\P~\t{of}~S}}
@@
 \begin{align}
-  N!_{\abs{S}} O_S
+  N!_{\abs{S}}\, O_S
   &\EQFS \sum_{\substack{\t{partitions}\\P~\t{of}~S}}
-  \p{-1}^{\abs{S}-\abs{P}} N^{\abs{P}} \O_P
-  \times \prod_{p\in P}\p{\abs{p}-1}!
+  \p{-1}^{\abs{S}-\abs{P}} \O_P \times \prod_{p\in P}\p{\abs{p}-1}!
   \label{eq:multi_to_collective_full} \\
@@
   \sum_{\substack{\t{ordered}\\\t{partitions}\\\tilde P~\t{of}~\Pi}}
-  \p{-1}^{\abs{S}-\abs{\tilde P}} N^{\abs{\tilde P}} \O_{\tilde P}
+  \p{-1}^{\abs{S}-\abs{\tilde P}} \O_{\tilde P}
   \times \prod_{\tilde p\in\tilde P} \p{\abs{\tilde p}-1}!
@@
 \begin{align}
-  N^{\abs{\tilde S}} \O_{\tilde S} \EQFS
+  \O_{\tilde S} \EQFS
   \sum_{\substack{\t{ordered}\\\t{partitions}\\\tilde P~\t{of}~\tilde S}}
@@
 \eqref{eq:multi_to_collective_full_simp} closely resemble each other,
-the latter can involve less computational overhead due to the lack of
-symmetrization in the construction of $\O_{\tilde P}$.  There are, in
-principle, even more economical decompositions of a multi-local
-operators into collective opreators, for example
+in practice the latter involves less computational overhead due to the
+lack of symmetrization in the construction of $\O_{\tilde P}$.  There
+are, in principle, even more economical decompositions of a
+multi-local operators into collective opreators, for example
 \begin{align}
   N!_2\, X\otimes Y
-  &\EQFS N^2\, \col{X}\,\col{Y} - N\, \col{XY}, \\
+  &\EQFS \col{X}\,\col{Y} - \col{XY}, \\
   N!_3\, X\otimes Y \otimes Z
-  &\EQFS N^3\, \col{X}\,\col{Y}\,\col{Z}
-  - N^2 \p{\col{X}\,\col{YZ} + \col{Y}\,\col{XZ} + \col{Z}\,\col{XY}}
-  + N \p{\col{ZXY} + \col{YXZ}},
+  &\EQFS \col{X}\,\col{Y}\,\col{Z}
+  - \col{X}\,\col{YZ} - \col{Y}\,\col{XZ} - \col{Z}\,\col{XY}
+  + \col{ZXY} + \col{YXZ},
 \end{align}
@@
 \begin{align}
-  N!_2\, X\otimes X \EQFS N^2\, {\col{X}\,}^2 - N\, \col{X^2},
-  &&
-  N!_3\, X^{\otimes 3} \EQFS N^3\, {\col{X}\,}^3
-  - 3 N^2\, \col{X} \odot \col{X^2} + 2 N\, \col{X^3},
+  N!_2\, X\otimes X \EQFS {\col{X}\,}^2 - \col{X^2},
+  &&
+  N!_3\, X^{\otimes 3} \EQFS {\col{X}\,}^3
+  - 3 \col{X} \odot \col{X^2} + 2\col{X^3},
 \end{align}
@@
   N!_4\, X^{\otimes 4}
-  \EQFS N^4\, {\col{X}\,}^4 - N^3 \p{4 {\col{X}\,}^2 \odot \col{X^2}
-    + 2 \col{X} \, {\col{X}\,}^2 \, \col{X}}
-  + N^3\p{3 {\col{X^2}\,}^2 + 8 \col{X} \odot \col{X^3}}
-  - 6 N\, \col{X^4}.
+  \EQFS {\col{X}\,}^4 - 4 {\col{X}\,}^2 \odot \col{X^2}
+  - 2 \col{X} \, {\col{X}\,}^2 \, \col{X}
+  + 3 {\col{X^2}\,}^2 + 8 \col{X} \odot \col{X^3}
+  - 6 \col{X^4}.
 \end{align}
@@
 \begin{align}
-  N!_2\, X\otimes X \EQFS N^2\, {\col{X}\,}^2 - N,
-  &&
-  N!_3\, X^{\otimes 3}
-  \EQFS N^3\, {\col{X}\,}^3 - \p{3N-2} N\, \col{X},
-\end{align}
-\begin{align}
-  N!_4\, X^{\otimes 4} \EQFS N^4\, {\col{X}\,}^4
-  - 2\p{3N-4} N^2\, {\col{X}\,}^2 + 3 N \p{N-2}.
+  N!_2\, X\otimes X \EQFS {\col{X}\,}^2 - N,
+  &&
+  N!_3\, X^{\otimes 3} \EQFS {\col{X}\,}^3 - \p{3N-2} \col{X},
+\end{align}
+\begin{align}
+  N!_4\, X^{\otimes 4} \EQFS {\col{X}\,}^4
+  - 2\p{3N-4} {\col{X}\,}^2 + 3 N \p{N-2}.
 \end{align}

### 2020-01-14 collective_sun/sun_perturbation.tex
restrict the two-body perturbation operators to those that obey pair-wise permutational symmetry

@@
 numbers; $X$ is a trace-zero single-body operator on $\H_n$; $O$ is a
-trace-zero two-body operator on $\H_n\otimes\H_n$; $\v v_X$ and $\v X$
-are $N$-component vectors defined by
+trace-zero two-body operator on $\H_n\otimes\H_n$ that obeys pair-wise
+permutational symmetry, i.e.~$O_{pq}=O_{qp}$; $\v v_X$ and $\v X$ are
+$N$-component vectors defined by
 \begin{align}

### 2020-01-16 collective_sun/sun_perturbation.tex
changed notation for denoting collective opetrators

@@
 \newcommand{\EQFS}{=_{\text{FS}}}
-\newcommand{\col}{\overline}
-
-\DeclareMathOperator{\sign}{sign}
-\DeclareMathOperator{\cov}{cov}
-\let\var\oldvar
-\DeclareMathOperator{\var}{var}
+\newcommand{\col}{\underline}
+
 \DeclareMathOperator{\diag}{diag}
-\let\flat\oldflat
-\DeclareMathOperator{\flat}{flat}
-\DeclareMathOperator{\counts}{counts}
-\newcommand{\flatvec}{\underline\flat\,}
 \newcommand{\ul}{\underline}
@@
   \f{\v v_X^\Delta\c\v v_Y^\Delta}{\Delta}
-  \p{\col{X}\,\col{Y} - N\, \col{XY}},
+  \p{\col{X}\,\col{Y} - N \col{XY}},
   \label{eq:H_1_2}
@@
   = - \f1{N\p{N-1}} \sum_{X,Y} \v v_X \c\v v_Y
-  \p{\col{X}\,\col{Y} - N\,\col{XY}}.
+  \p{\col{X}\,\col{Y} - N \col{XY}}.
 \end{align}
@@
 \begin{align}
-  \ul{\check{\m h}}
+  \ut{\check{\m h}}
   \equiv \sum_{d\ne0} h_{d,0} \op{d}
@@
   &&
-  \ul{\v{\m w}} \equiv \sum_{d\ne0} w_{d,0} \ket{d},
+  \ut{\v{\m w}} \equiv \sum_{d\ne0} w_{d,0} \ket{d},
 \end{align}
@@
 \begin{align}
-  \ul{\check{\m h}} \c \ul{\v{\m w}}
-  = \p{\f12\Delta + h_0} \ul{\v{\m w}}.
+  \ut{\check{\m h}} \c \ut{\v{\m w}}
+  = \p{\f12\Delta + h_0} \ut{\v{\m w}}.
   \label{eq:cond_2_square}
@@
 projector onto the reflection-symmetric subspace and expand
-$\ul{\v{\m w}}$ as
+$\ut{\v{\m w}}$ as
 \begin{align}
@@
   &&
-  \ul{\v{\m w}} = \sum_{d\in\VV_N} w_{d,0} \,
+  \ut{\v{\m w}} = \sum_{d\in\VV_N} w_{d,0} \,
   \eta_d \sqrt{2} \ket{d_+}.
 \end{align}
-Finally, we project $\ul{\check{\m h}}$ onto the reflection-symmetric
+Finally, we project $\ut{\check{\m h}}$ onto the reflection-symmetric
 subspace,
 \begin{align}
-  \P_+ \ul{\check{\m h}} \P_+
+  \P_+ \ut{\check{\m h}} \P_+
   &= \sum_{a,b\in\VV_N} \sum_{c,d\ne0}
@@
 \begin{align}
-  \P_+ \ul{\check{\m h}} \P_+ \c \ul{\v{\m w}}
-  = \p{\f12\Delta + h_0} \ul{\v{\m w}}.
+  \P_+ \ut{\check{\m h}} \P_+ \c \ut{\v{\m w}}
+  = \p{\f12\Delta + h_0} \ut{\v{\m w}}.
 \end{align}
 Note that the factors of $\eta_a,\eta_b$ in
-$\P_+ \ul{\check{\m h}} \P_+$ only matter when $N$ is even, in which
+$\P_+ \ut{\check{\m h}} \P_+$ only matter when $N$ is even, in which
 case these factors merely multiply the last row and last column of
-$\P_+ \ul{\check{\m h}} \P_+$ by a factor of $1/\sqrt{2}$.
+$\P_+ \ut{\check{\m h}} \P_+$ by a factor of $1/\sqrt{2}$.
 
@@
 \end{align}
-where we place a bar over a single-spin operator to denote its
-collective version, i.e.~$\col{Q}\equiv\sum_j Q^{(j)}$, where
-$Q^{(j)}$ is an operator that acts with the single-spin operator $Q$
-on spin $j$ and trivially on all other spins.  By construction, the
-set of collective operators $\set{\O_P:\t{partitions}~P~\t{of}~S}$ is
-one-to-one to the set of multi-local operators
-$\set{O_P:\t{partitions}~P~\t{of}~S}$, and it is straightforward to
-work out that for any set $S$ of single-spin operators,
+where we underline a single-spin operator to denote its collective
+version, i.e.~$\col{Q}\equiv\sum_j Q^{(j)}$, where $Q^{(j)}$ is an
+operator that acts with the single-spin operator $Q$ on spin $j$ and
+trivially on all other spins.  By construction, the set of collective
+operators $\set{\O_P:\t{partitions}~P~\t{of}~S}$ is one-to-one to the
+set of multi-local operators $\set{O_P:\t{partitions}~P~\t{of}~S}$,
+and it is straightforward to work out that for any set $S$ of
+single-spin operators,
 \begin{align}
@@
   \EQFS \sum_{\substack{\t{partitions}\\P~\t{of}~S}}
-  N!_{\abs{P}}\, O_P,
+  N!_{\abs{P}} O_P,
   &&
@@
 \begin{align}
-  N!_{\abs{S}}\, O_S
+  N!_{\abs{S}} O_S
   &\EQFS \sum_{\substack{\t{partitions}\\P~\t{of}~S}}
@@
 \begin{align}
-  N!_2\, X\otimes Y
+  N!_2 X\otimes Y
   &\EQFS \col{X}\,\col{Y} - \col{XY}, \\
-  N!_3\, X\otimes Y \otimes Z
+  N!_3 X\otimes Y \otimes Z
   &\EQFS \col{X}\,\col{Y}\,\col{Z}
@@
 \begin{align}
-  N!_2\, X\otimes X \EQFS {\col{X}\,}^2 - \col{X^2},
-  &&
-  N!_3\, X^{\otimes 3} \EQFS {\col{X}\,}^3
+  N!_2 X\otimes X \EQFS \col{X}^2 - \col{X^2},
+  &&
+  N!_3 X^{\otimes 3} \EQFS \col{X}^3
   - 3 \col{X} \odot \col{X^2} + 2\col{X^3},
@@
 \begin{align}
-  N!_4\, X^{\otimes 4}
-  \EQFS {\col{X}\,}^4 - 4 {\col{X}\,}^2 \odot \col{X^2}
-  - 2 \col{X} \, {\col{X}\,}^2 \, \col{X}
-  + 3 {\col{X^2}\,}^2 + 8 \col{X} \odot \col{X^3}
+  N!_4 X^{\otimes 4}
+  \EQFS \col{X}^4 - 4 \col{X}^2 \odot \col{X^2}
+  - 2 \col{X} \, \col{X}^2 \, \col{X}
+  + 3 \p{\col{X^2}}^2 + 8 \col{X} \odot \col{X^3}
   - 6 \col{X^4}.
@@
 \begin{align}
-  N!_2\, X\otimes X \EQFS {\col{X}\,}^2 - N,
-  &&
-  N!_3\, X^{\otimes 3} \EQFS {\col{X}\,}^3 - \p{3N-2} \col{X},
-\end{align}
-\begin{align}
-  N!_4\, X^{\otimes 4} \EQFS {\col{X}\,}^4
-  - 2\p{3N-4} {\col{X}\,}^2 + 3 N \p{N-2}.
+  N!_2 X\otimes X \EQFS \col{X}^2 - N,
+  &&
+  N!_3 X^{\otimes 3} \EQFS \col{X}^3 - \p{3N-2} \col{X},
+\end{align}
+\begin{align}
+  N!_4 X^{\otimes 4} \EQFS \col{X}^4
+  - 2\p{3N-4} \col{X}^2 + 3 N \p{N-2}.
 \end{align}

### 2020-01-16 collective_sun/sun_perturbation.tex
address the case of a translationally invariant lattice with dimension D>1

@@
 \end{align}
-The eigenvalue problem in \eqref{eq:cond_2} can be simplified somewhat
+The eigenvalue problem in \eqref{eq:cond_2} can be simplified further
 when the interaction Hamiltonian $H_0$ is translationally invariant;
@@
 the two-body eigenvalue problem to that of diagonalizing an
-$\floor{N/2}\times\floor{N/2}$ matrix.
+$\p{N-1}\times\p{N-1}$ matrix.  Finally, we discuss the reduction of
+the two-body problem to that of diagonalizing a
+$\sim\p{N/2^D}\times\p{N/2^D}$ matrix, and work out the case of $D=1$
+explicitly as an example.
 
@@
 \end{align}
-where $k\in\ZZ_N\times2\pi/N$.  The corresponding eigenvalues can be
-determined by expanding
+where $k\in\ZZ_N\times2\pi/N$ (with
+$\ZZ_N\equiv\set{0,1,\cdots,N-1}$).  The corresponding eigenvalues can
+be determined by expanding
 \begin{align}
@@
   = \sum_{p,d} h_{p,d+p} e^{i\p{d+p}k} \ket{p}
-  = \sum_d h_{0,d} e^{idk} \sum_p e^{ipk}\ket{p}
-  = \sum_d h_{0,d} \cos\p{dk} \v v_k.
-\end{align}
-The remainder of \eqref{eq:cond_1_ref} that we need to sort out is
-$\diag\v h$, where all $h_p=\sum_qh_{pq}=\sum_dh_{0,d}$ are equal,
-which implies that $\diag\v h=\sum_dh_{0,d}$ is a scalar.  We thus
-find that
+  = \sum_d h_{0,d} e^{idk} \v v_k
+  = \sum_d h_{0,d} \cos\p{dk} \v v_k,
+\end{align}
+where the imaginary terms vanish due to the fact that
+$h_{0,d}=h_{0,-d}$.  The remainder of \eqref{eq:cond_1_ref} that we
+need to sort out is $\diag\v h$, where all
+$h_p=\sum_qh_{pq}=\sum_dh_{0,d}$ are equal, which implies that
+$\diag\v h=\sum_dh_{0,d}$ is a scalar.  We thus find that
 \begin{align}
@@
 \end{align}
-Note that $\Delta_k=\Delta_{-k}$, which implies that there are at most
-$\lceil N/2\rceil$ distinct eigenvalues of $\m h - \diag\v h$ (one of
-which is $\Delta_0=0$).
+This solution also holds for a $D$-dimensional periodic lattice of
+$N=L^D$ spins, with $d\in\ZZ_L^D$ and $k\in\ZZ_L^D\times 2\pi/L$ (and
+$dk\to d\c k\equiv\sum_{j=1}^D d_j k_j$).
 
@@
 and perturbation $\V_2\p{\m w}$ are translationally invariant, then
-the couplings $h_{pq}$ and coefficients $w_{pq}$ depend only on the
-separation $\abs{p-q}$.  In this case, all $h_p=h_0$ are equal, and
-\eqref{eq:cond_2_ref_start} can be written in the form
+the couplings $h_{pq}$ and coefficients $w_{pq}$ depend on $p$ and $q$
+only through the separation $\abs{p-q}$.  In this case, all $h_p=h_0$
+are equal, and \eqref{eq:cond_2_ref_start} can be written in the form
 \begin{align}
@@
 \end{align}
-We can reduce this eigenvalue problem even further by making use of a
-reflection symmetry that comes from the fact that $h_{pq}=h_{qp}$ and
-$w_{pq}=w_{qp}$, which implies $h_{d,0}=h_{-d,0}$ and
-$w_{d,0}=w_{-d,0}$ (where all indices are taken modulo $N$, taking
-$-d\to N-d$).  To this end, we define a complete basis of orthogonal
-vectors that are symmetric under reflection, namely for all
-$d\in\VV_N\equiv\set{1,2,\cdots,\floor{N/2}}$,
+This eigenvalue problem takes an identical form a $D$ dimensional
+periodic lattice of $N=L^D$ spins, in which case lattice sites and
+shift matrices are indexed by vectors $d\in\ZZ_L^D$.
+
+We can reduce the eigenvalue problem in \eqref{eq:cond_2_square} even
+further by making use of reflection symmetries such as
+$h_{d,0}=h_{-d,0}$.  In $D$ dimensions, there is one such symmetry for
+each coordinate that indexes lattice sites, which implies that the
+degrees of freedom in $\ut{\v{\m w}}$ gets reduced by a factor of
+$\sim 2^D$.  Reflection symmetries similarly reduce the degrees of
+freedom in $\ut{\check{\m h}}$ from $\sim N\times N$ to
+$\sim\p{N/2^D}\times\p{N/2^D}$.  We work out the case of $D=1$ below
+as an example, and leave the general case of arbitrary dimension
+$D\ge1$ to the future, if and when it becomes necessary.
+
+In order to exploit reflection symmetry in one dimension ($D=1$), we
+define a complete basis of orthogonal vectors that are symmetric under
+reflection: for all $d\in\VV_N\equiv\set{1,2,\cdots,\floor{N/2}}$,
 \begin{align}

### 2020-01-16 collective_sun/sun_perturbation.tex
some improvements to text, removing the implicit assumption of dimension D=1 by default

@@
 \end{align}
-where $\diag\v h\equiv\sum_p h_p \op{p}$ is a matrix with $\v h$ on
-the diagonal and zeros everywhere else.  Solving the general
-eigenvalue problem in \eqref{eq:cond_1} requires diagonalizing the
-$N\times N$ matrix $\m h-\diag\v h$.  When the interaction Hamiltonian
-$H_0$ is translationally invariant, however, this eigenvalue problem
-can be solved analytically; we provide the corresponding solution in
-Appendix \ref{sec:trans_inv}.
+where $\m h$ is a matrix of all $h_{pq}$, and
+$\diag\v h\equiv\sum_p h_p \op{p}$ is a matrix with $\v h$ on the
+diagonal and zeros everywhere else.  Solving the general eigenvalue
+problem in \eqref{eq:cond_1} requires diagonalizing the $N\times N$
+matrix $\m h-\diag\v h$.  When the interaction Hamiltonian $H_0$ is
+translationally invariant, however, this eigenvalue problem can be
+solved analytically; we provide the corresponding solution in Appendix
+\ref{sec:trans_inv}.
 
@@
 \end{align}
-The eigenvalue problem in \eqref{eq:cond_2} can be simplified further
-when the interaction Hamiltonian $H_0$ is translationally invariant;
-we discuss this simplification in Appendix \ref{sec:trans_inv}.
+Solving the general eigenvalue problem in \eqref{eq:cond_2} requires
+diagonalizing the ${N \choose 2}\times{N \choose 2}\sim N^2\times N^2$
+matrix $\check{\m h}$.  When the interaction Hamiltonian $H_0$ is
+translationally invariant, however, this eigenvalue problem can be
+reduced to that of diagonalizing an $\sim N\times N$ matrix; we
+discuss this reduction in Appendix \ref{sec:trans_inv}.
 
@@
 the two-body eigenvalue problem to that of diagonalizing an
-$\p{N-1}\times\p{N-1}$ matrix.  Finally, we discuss the reduction of
-the two-body problem to that of diagonalizing a
-$\sim\p{N/2^D}\times\p{N/2^D}$ matrix, and work out the case of $D=1$
+$\p{N-1}\times\p{N-1}$ matrix.  Finally, we discuss the further
+reduction of the two-body problem to that of diagonalizing a matrix
+with dimensions $\sim\p{N/2^D}\times\p{N/2^D}$ in the case of
+$D$-dimensional periodic lattice, and work out the case of $D=1$
 explicitly as an example.
@@
 \begin{align}
-  \v v_k \equiv \sum_p e^{ipk} \ket{p},
-\end{align}
-where $k\in\ZZ_N\times2\pi/N$ (with
-$\ZZ_N\equiv\set{0,1,\cdots,N-1}$).  The corresponding eigenvalues can
-be determined by expanding
-\begin{align}
-  \m h\c\v v_k = \sum_{p,q} h_{pq} e^{iqk} \ket{p}
-  = \sum_{p,d} h_{p,d+p} e^{i\p{d+p}k} \ket{p}
-  = \sum_d h_{0,d} e^{idk} \v v_k
-  = \sum_d h_{0,d} \cos\p{dk} \v v_k,
+  \v v_k \equiv \sum_p e^{ip\c k} \ket{p},
+\end{align}
+where on a $D$-dimensional periodic lattice of $N=L^D$ spins, lattice
+sites are indexed by vectors $p\in\ZZ_L^D$, and wavenumbers take on
+values $k\in\ZZ_L^D\times2\pi/N$.  The corresponding eigenvalues of
+$\m h$ can be determined by expanding
+\begin{align}
+  \m h\c\v v_k = \sum_{p,q} h_{pq} e^{iq\c k} \ket{p}
+  = \sum_{p,d} h_{p,d+p} e^{i\p{d+p}\c k} \ket{p}
+  = \sum_d h_{0,d} e^{id\c k} \v v_k
+  = \sum_d h_{0,d} \cos\p{d\c k} \v v_k,
 \end{align}
@@
   &&
-  \Delta_k \equiv \sum_d h_{0,d} \sp{\cos\p{dk}-1}.
-\end{align}
-This solution also holds for a $D$-dimensional periodic lattice of
-$N=L^D$ spins, with $d\in\ZZ_L^D$ and $k\in\ZZ_L^D\times 2\pi/L$ (and
-$dk\to d\c k\equiv\sum_{j=1}^D d_j k_j$).
+  \Delta_k \equiv \sum_d h_{0,d} \sp{\cos\p{d\c k}-1}.
+\end{align}
 
@@
 \end{align}
-where $\m h\circ\m w$ denotes an element-wise (Kronecker) product of
-$\m h$ and $\m w$, i.e.~with matrix elements $h_{pq}w_{pq}$.  In order
-to find solutions to \eqref{eq:cond_2_ref}, we we expand the matrices
+where $\m h$ and $\m w$ are matrices of all $h_{pq}$ and $w_{pq}$, and
+$\m h\circ\m w$ denotes an element-wise (Kronecker) product of $\m h$
+and $\m w$, i.e.~with matrix elements $h_{pq}w_{pq}$.  In order to
+find solutions to \eqref{eq:cond_2_ref}, we we expand the matrices
 $\m h,\m w$ in a basis that respects translational symmetry.  To this
-end, for all $d\in\ZZ_N$ we define the {\it shift matrices}
+end, for all lattice site displacements $d\in\ZZ_L^D$ we define the
+{\it shift matrices}
 \begin{align}
@@
 \end{align}
-where the sum $p+d$ is implicitly taken modulo $N$.  The shift
+where the element-wise sums in $p+d$ are implicitly taken modulo $L$
+(i.e.~the length of the periodic lattice along each axis).  The shift
 matrices allow us to expand
@@
 \end{align}
-This eigenvalue problem takes an identical form a $D$ dimensional
-periodic lattice of $N=L^D$ spins, in which case lattice sites and
-shift matrices are indexed by vectors $d\in\ZZ_L^D$.
-
-We can reduce the eigenvalue problem in \eqref{eq:cond_2_square} even
-further by making use of reflection symmetries such as
-$h_{d,0}=h_{-d,0}$.  In $D$ dimensions, there is one such symmetry for
-each coordinate that indexes lattice sites, which implies that the
-degrees of freedom in $\ut{\v{\m w}}$ gets reduced by a factor of
+We can reduce this eigenvalue problem even further by making use of
+reflection symmetries such as $h_{d,0}=h_{-d,0}$.  In $D$ dimensions,
+there is one such symmetry for each lattice axis, which implies that
+the degrees of freedom in $\ut{\v{\m w}}$ gets reduced by a factor of
 $\sim 2^D$.  Reflection symmetries similarly reduce the degrees of
@@
 as an example, and leave the general case of arbitrary dimension
-$D\ge1$ to the future, if and when it becomes necessary.
+$D\ge1$ to future work.
 

### 2020-01-17 collective_sun/sun_perturbation.tex
fixed typo in Eq. (25) added a proof (new appendix) of a non-vanishing spectral gap for power-law interactions with \alpha <= D

@@
 
-\newcommand{\EE}{\mathbb{E}}
-\newcommand{\RR}{\mathbb{R}}
 \renewcommand{\SS}{\mathbb{S}}
+\newcommand{\TT}{\mathbb{T}}
 \newcommand{\VV}{\mathbb{V}}
@@
 states by an interaction energy difference that is large compared to
-any coupling between $\M_0$ and its orthogonal complement.  The
-precise conditions for such a non-vanishing many-body interaction
-energy gap is an important question that deserves further study in
-future work.
+any coupling between $\M_0$ and its orthogonal complement.  Power-law
+couplings of the form $h_{pq}=-h/\abs{p-q}^\alpha$ on a
+$D$-dimensional lattice, for example, yield a non-vanishing spectral
+gap in the thermodynamic limit when $\alpha\le D$ (see Appendix
+\ref{sec:gap}).
 
@@
 \begin{align}
-  H_2^{(2)} = - \P_0 \V_2^\Delta \E \V_2^\Delta \P_0
+  H_2^{(2)} = - \P_0 \V_2 \E \V_2 \P_0
   = -\sum_{\Delta\ne0} \f1\Delta \P_0 \p{\V_2^\Delta}^2 \P_0,
@@
 where $s_\z\equiv\p{\op\up-\op\dn}/2$ is an SU(2) spin-$z$ operator,
-then
+the second-order effective Hamiltonian $H_2^{(2)}$ takes the form
 \begin{align}
@@
 $\check{\m h}$ defined in \eqref{eq:h_super_mat} by $\v{\m m}_\Delta$,
-which allows us to expand
+which allows us to concisely write
 \begin{align}
@@
 \begin{align}
-  N!_k \equiv \prod_{j=0}^{k-1} \p{N-j} = \f{N!}{\p{N-k}!}.
-\end{align}
+  N!_k \equiv \prod_{j=0}^{k-1} \p{N-j} = \f{N!}{\p{N-k}!}
+\end{align}
+denotes a falling factorial.
 
@@
 \appendix
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Existence of a spectral gap with power-law interactions}
+\label{sec:gap}
+
+Here we show that power-law SU($n$)-symmetric interactions of the form
+in \eqref{eq:ints} with $h_{pq}=-h/\abs{p-q}^\alpha$ yield a
+non-vanishing many-body interaction energy gap when $\alpha<D$, where
+$D$ is the dimension of the lattice.  On a periodic lattice of $N=L^D$
+spins, the spectral gap of the interaction Hamiltonian is (see
+Appendix \ref{sec:trans_inv_single})
+\begin{align}
+  \Delta_D
+  = \sum_{d\in\tilde\ZZ_L^D} h_{0,d} \sp{\cos\p{d \c k_{\t{SE}}}-1}
+  = h \sum_{\substack{d\in\tilde\ZZ_L^D\\d\ne0}}
+  \f{1-\cos\p{d \c k_{\t{SE}}}}{\abs{d}^\alpha},
+\end{align}
+where the domain of $d$ is given by
+$\tilde\ZZ_L\equiv\set{-\floor{L/2},-\floor{L/2}+1,\cdots,\ceil{L/2}}$,
+and $k_{\t{SE}}$ is a wavenumber for a singly-excited spin-wave state.
+In order to yield the smallest excitation energy $\Delta_D$, the
+wavenumber $k_{\t{SE}}$ should maximize the contribution of the cosine
+term above, which is achieved by a wavenumber that minimizes the
+oscillations of this term when integrated over the entire lattice.  A
+suitable candidate for a minimal wavenumber is
+$k_{\t{SE}}=\p{2\pi/L,0,0,\cdots}$, which corresponds to an excitation
+energy
+\begin{align}
+  \Delta_D = h \sum_{\substack{d\in\tilde\ZZ_L^D\\d\ne0}}
+  \f{1-\cos\p{d_12\pi/L}}{\abs{d}^\alpha}.
+\end{align}
+Defining $\epsilon\equiv2/L$ and a rescaled domain
+$\SS_\epsilon\equiv\tilde\ZZ_L/\epsilon\subset\sp{-1,1}$, we
+substitute $x\equiv\epsilon d\in\SS_\epsilon^D$ to get
+\begin{align}
+  \Delta_D
+  = h \sum_{\substack{x\in\SS_\epsilon^D\\x\ne0}}
+  \f{1-\cos\p{\pi x_1}}{\abs{x/\epsilon}^\alpha}
+  = h \epsilon^{D-\alpha} \sum_{\substack{x\in\SS_L^D\\x\ne0}}
+  \epsilon^D \f{1-\cos\p{\pi x_1}}{\abs{x}^\alpha}.
+\end{align}
+As $\epsilon\to0$ the discrete sum over $x$ becomes an integral that
+avoids an infinitesimal region about the origin, i.e.
+\begin{align}
+  \Delta_D = h \epsilon^{\alpha-D} \I_{D\epsilon},
+  &&
+  \I_{D\epsilon}
+  \equiv \int_{\TT_1^D\setminus\TT_\epsilon^D} \d^Dx\,
+  \f{1-\cos\p{\pi x_1}}{\abs{x}^\alpha},
+\end{align}
+where the integral $\I_{D\epsilon}$ is defined using the interval
+$\TT_a\equiv\p{-a,a}$.  The integrand of $\I_{D\epsilon}$ is strictly
+positive and well-behaved on the entirety of its domain except for the
+origin, where depending on the value of $\alpha$ the integrand may
+vanish or diverge as $\abs{x}\to0$.  Together, these facts mean that
+\begin{align}
+  \I_{D\epsilon} \stackrel{\epsilon\to0}{\sim} \epsilon^{-\gamma},
+  &&
+  \Delta_D \stackrel{\epsilon\to0}{\sim} h \epsilon^{\alpha-D-\gamma},
+\end{align}
+for some $\gamma\ge0$, which implies that gap $\Delta_D$ is
+non-vanishing when
+\begin{align}
+  \alpha \le D + \gamma,
+\end{align}
+and in particular when $\alpha \le D$.
 
@@
 where on a $D$-dimensional periodic lattice of $N=L^D$ spins, lattice
-sites are indexed by vectors $p\in\ZZ_L^D$, and wavenumbers take on
-values $k\in\ZZ_L^D\times2\pi/N$.  The corresponding eigenvalues of
-$\m h$ can be determined by expanding
+sites are indexed by vectors
+$p\in\tilde\ZZ_L^D \equiv
+\set{-\floor{L/2},-\floor{L/2}+1,\cdots,\ceil{L/2}}$, and wavenumbers
+take on values $k\in\tilde\ZZ_L^D\times2\pi/L$.  The corresponding
+eigenvalues of $\m h$ can be determined by expanding
 \begin{align}
@@
 $\m h,\m w$ in a basis that respects translational symmetry.  To this
-end, for all lattice site displacements $d\in\ZZ_L^D$ we define the
-{\it shift matrices}
+end, for all lattice site displacements $d\in\tilde\ZZ_L^D$ we define
+the {\it shift matrices}
 \begin{align}

### 2020-01-17 collective_sun/sun_perturbation.tex
added proof that the gap vanishes when \alpha > D

@@
 \newcommand{\I}{\mathcal{I}}
+\newcommand{\K}{\mathcal{K}}
 \renewcommand{\L}{\mathcal{L}}
@@
 for some $\gamma\ge0$, which implies that gap $\Delta_D$ is
-non-vanishing when
-\begin{align}
-  \alpha \le D + \gamma,
-\end{align}
-and in particular when $\alpha \le D$.
+non-vanishing when $\alpha\le D\le D+\gamma$.
+
+For the sake of completion, we add that in fact
+$\gamma=\max\set{0,\alpha-D-2}$, which implies that the gap $\Delta_D$
+vanishes in the thermodynamic limit when $\alpha>D$.  The value of
+$\gamma$ is determined by the behavior of the integrand of
+$\I_{D\epsilon}$ near $\abs{x}\sim\epsilon\ll1$, where
+$1-\cos\p{\pi x_1}\sim x_1^2$, so
+\begin{align}
+  \I_{D\epsilon}
+  \stackrel{\epsilon\to0}{\sim} \K_{D\epsilon}
+  \equiv \int_{\TT_1^D\setminus\TT_\epsilon^D} \d^Dx\,
+  \f{x_1^2}{\abs{x}^\alpha}
+  = \int_{\TT_1^D\setminus\TT_\epsilon^D} \d y\, \d z^{D-1}
+  \f{y^2}{\p{y^2+z^2}^{\alpha/2}},
+\end{align}
+where for convenience we introduced the variables $y\equiv x_1$ and
+$z\equiv\p{x_2,x_3,\cdots,x_D}$.  The integral $\K_{D\epsilon}$ can be
+split up into two contributions: a ``bulk'' term
+$\K_{D\epsilon}^{\t{B}}$ that integrates over
+$\abs{z}\gtrsim\epsilon$, and a remainder $\K_{D\epsilon}^{\t{R}}$
+that integrates over $\abs{z}\lesssim\epsilon$.  The domain of $y$ in
+each case is determined by the restriction that
+$\p{y,z}\notin\TT_\epsilon^D$.  Altogether,
+\begin{align}
+  \K_{D\epsilon} = \K_{D\epsilon}^{\t{B}} + \K_{D\epsilon}^{\t{R}},
+\end{align}
+where
+\begin{align}
+  \K_{D\epsilon}^{\t{B}}
+  &\equiv \int_{\TT_1} \d y
+  \int_{\TT_1^{D-1}\setminus\TT_\epsilon^{D-1}} \d^{D-1}z\,
+  \f{y^2}{\p{y^2+z^2}^{\alpha/2}}, \\
+  \K_{D\epsilon}^{\t{R}}
+  &\equiv \int_{\TT_1\setminus\TT_\epsilon} \d y
+  \int_{\TT_\epsilon^{D-1}} \d^{D-1}z\,
+  \f{y^2}{\p{y^2+z^2}^{\alpha/2}}.
+\end{align}
+Changing to spherical coordinates, we can say that
+\begin{align}
+  \K_{D\epsilon}^{\t{B}}
+  \sim \int_0^1 \d y \int_\epsilon^1 \d z\,
+  \f{y^2 z^{D-2}}{\p{y^2+z^2}^{\alpha/2}},
+  &&
+  \K_{D\epsilon}^{\t{R}}
+  \sim \int_\epsilon^1 \d y \int_0^\epsilon \d z\,
+  \f{y^2 z^{D-2}}{\p{y^2+z^2}^{\alpha/2}},
+\end{align}
+where the resulting integrals can be evaluated on a computer algebra
+system (e.g.~Mathematica) and expanded to leading orders in $\epsilon$
+to find that
+\begin{align}
+  \K_{D\epsilon}^{\t{B}} \stackrel{\epsilon\to0}{\sim}
+  B_0 + B_1 \epsilon^{D-\alpha+2},
+  &&
+  \K_{D\epsilon}^{\t{R}} \stackrel{\epsilon\to0}{\sim}
+  R_1 \epsilon^{D-\alpha+2},
+\end{align}
+for some constants $B_0,B_1,R_1$ that generally depend on $D$ and
+$\alpha$.  It follows that
+\begin{align}
+  \I_{H\epsilon} \stackrel{\epsilon\to0}{\sim} \K_{D\epsilon}
+  = \K_{D\epsilon}^{\t{B}} + \K_{D\epsilon}^{\t{R}}
+  \stackrel{\epsilon\to0}{\sim} \epsilon^{-\gamma},
+  &&
+  \gamma = -\min\set{0,D-\alpha+2}
+\end{align}
+and so
+\begin{align}
+  \Delta_D
+  \stackrel{\epsilon\to0}{\sim} \epsilon^{\alpha-D-\gamma}
+  = \epsilon^{\min\set{\alpha-D,2}}.
+\end{align}
 

### 2020-01-17 collective_sun/sun_perturbation.tex
simplify proof of a vanishing gap when \alpha > D

--- 2020-01-17 collective_sun/sun_perturbation.tex
@@
 in \eqref{eq:ints} with $h_{pq}=-h/\abs{p-q}^\alpha$ yield a
-non-vanishing many-body interaction energy gap when $\alpha<D$, where
-$D$ is the dimension of the lattice.  On a periodic lattice of $N=L^D$
-spins, the spectral gap of the interaction Hamiltonian is (see
+non-vanishing many-body interaction energy gap when $\alpha\le D$,
+where $D$ is the dimension of the lattice.  On a periodic lattice of
+$N=L^D$ spins, the spectral gap of the interaction Hamiltonian is (see
 Appendix \ref{sec:trans_inv_single})
@@
 for some $\gamma\ge0$, which implies that gap $\Delta_D$ is
-non-vanishing when
-\begin{align}
-  \alpha \le D + \gamma,
-\end{align}
-and in particular when $\alpha \le D$.
+non-vanishing when $\alpha\le D\le D+\gamma$.
+
+For the sake of completion, we add that in fact the gap $\Delta_D$
+always vanishes in the thermodynamic limit when $\alpha>D$.  To see
+this behavior, we note that the asymptotic dependence of the integral
+$I_{D\epsilon}$ on $\epsilon$ is determined by the behavior of its
+integrand when $\abs{x}\sim\epsilon$, in which case
+$1-\cos\p{\pi x_1}\sim x_1^2$, so
+\begin{align}
+  \I_{D\epsilon}
+  \sim \int_{\TT_1^D\setminus\TT_\epsilon^D} \d^Dx\,
+  \f{x_1^2}{\abs{x}^\alpha}.
+\end{align}
+We can then use the fact that $x_1^2\le\abs{x}^2$ and change to
+spherical coordinates to find that
+\begin{align}
+  \I_{D\epsilon} \lesssim
+  \int_{\TT_1^D\setminus\TT_\epsilon^D} \d^Dx\,
+  \f{\abs{x}^2}{\abs{x}^\alpha}
+  \sim \int_\epsilon^1 \d x\, x^{D+1-\alpha}
+  \sim
+  \begin{cases}
+    \epsilon^0 & \alpha < D+2 \\
+    \log\p{1/\epsilon} & \alpha = D+2 \\
+    \epsilon^{D+2-\alpha} & \alpha > D+2
+  \end{cases}.
+\end{align}
+It follows that the gap
+\begin{align}
+  \Delta_D \stackrel{\epsilon\to0}{\lesssim}
+  \begin{cases}
+    \epsilon^{\alpha-D} & \alpha < D+2 \\
+    \epsilon^2 \log\p{1/\epsilon} & \alpha = D + 2 \\
+    \epsilon^2 & \alpha > D+2
+  \end{cases},
+\end{align}
+which vanishes as $\epsilon\to0$ for all $\alpha>D$.
 

### 2020-01-17 collective_sun/sun_perturbation.tex
minor change in notation

@@
 \begin{align}
-  \Delta_D
+  \Delta
   = \sum_{d\in\tilde\ZZ_L^D} h_{0,d} \sp{\cos\p{d \c k_{\t{SE}}}-1}
@@
 and $k_{\t{SE}}$ is a wavenumber for a singly-excited spin-wave state.
-In order to yield the smallest excitation energy $\Delta_D$, the
+In order to yield the smallest excitation energy $\Delta$, the
 wavenumber $k_{\t{SE}}$ should maximize the contribution of the cosine
@@
 \begin{align}
-  \Delta_D = h \sum_{\substack{d\in\tilde\ZZ_L^D\\d\ne0}}
+  \Delta = h \sum_{\substack{d\in\tilde\ZZ_L^D\\d\ne0}}
   \f{1-\cos\p{d_12\pi/L}}{\abs{d}^\alpha}.
@@
 \begin{align}
-  \Delta_D
+  \Delta
   = h \sum_{\substack{x\in\SS_\epsilon^D\\x\ne0}}
@@
 \begin{align}
-  \Delta_D = h \epsilon^{\alpha-D} \I_{D\epsilon},
+  \Delta = h \epsilon^{\alpha-D} \I_{D\epsilon},
   &&
@@
   &&
-  \Delta_D \stackrel{\epsilon\to0}{\sim} h \epsilon^{\alpha-D-\gamma},
-\end{align}
-for some $\gamma\ge0$, which implies that gap $\Delta_D$ is
+  \Delta \stackrel{\epsilon\to0}{\sim} h \epsilon^{\alpha-D-\gamma},
+\end{align}
+for some $\gamma\ge0$, which implies that gap $\Delta$ is
 non-vanishing when $\alpha\le D\le D+\gamma$.
 
-For the sake of completion, we add that in fact the gap $\Delta_D$
+For the sake of completion, we add that in fact the gap $\Delta$
 always vanishes in the thermodynamic limit when $\alpha>D$.  To see
@@
 \end{align}
-It follows that the gap
-\begin{align}
-  \Delta_D \stackrel{\epsilon\to0}{\lesssim}
+It follows that the spectral gap
+\begin{align}
+  \Delta \stackrel{\epsilon\to0}{\lesssim}
   \begin{cases}

### 2020-01-17 collective_sun/sun_perturbation.tex
clarified the origin of the relection symmetry in Appendix D; reflection about the origin is *always* a symmetry, whereas other reflection symmetries only occur for isotropic systems

@@
 the two-body eigenvalue problem to that of diagonalizing an
-$\p{N-1}\times\p{N-1}$ matrix.  Finally, we discuss the further
+$\p{N-1}\times\p{N-1}$ matrix.  Finally, for we discuss the further
 reduction of the two-body problem to that of diagonalizing a matrix
-with dimensions $\sim\p{N/2^D}\times\p{N/2^D}$ in the case of
-$D$-dimensional periodic lattice, and work out the case of $D=1$
-explicitly as an example.
+with dimensions $\sim\p{N/2^D}\times\p{N/2^D}$ in the case of an
+isotropic system on a $D$-dimensional periodic lattice, and we work
+out the case of $D=1$ explicitly as an example.
 
@@
 interaction Hamiltonian $H_0$ is translationally invariant, the
-couplings $h_{pq}$ depend only on the separation $\abs{p-q}$, so the
+couplings $h_{pq}$ depend only on the separation $\pm\p{p-q}$, so the
 eigenvectors of $\m h$ are plane waves of the form
@@
   \m h\c\v v_k = \sum_{p,q} h_{pq} e^{iq\c k} \ket{p}
-  = \sum_{p,d} h_{p,d+p} e^{i\p{d+p}\c k} \ket{p}
+  = \sum_{p,d} h_{p,p+d} e^{i\p{p+d}\c k} \ket{p}
   = \sum_d h_{0,d} e^{id\c k} \v v_k
@@
 \end{align}
-where the imaginary terms vanish due to the fact that
-$h_{0,d}=h_{0,-d}$.  The remainder of \eqref{eq:cond_1_ref} that we
-need to sort out is $\diag\v h$, where all
-$h_p=\sum_qh_{pq}=\sum_dh_{0,d}$ are equal, which implies that
-$\diag\v h=\sum_dh_{0,d}$ is a scalar.  We thus find that
+where the imaginary terms vanish because $h_{0,d}=h_{0,-d}$.  The
+remainder of \eqref{eq:cond_1_ref} that we need to sort out is
+$\diag\v h$, where all $h_p=\sum_qh_{pq}=\sum_dh_{0,d}$ are equal,
+which implies that $\diag\v h=\sum_dh_{0,d}$ is a scalar.  We thus
+find that
 \begin{align}
@@
 the couplings $h_{pq}$ and coefficients $w_{pq}$ depend on $p$ and $q$
-only through the separation $\abs{p-q}$.  In this case, all $h_p=h_0$
+only through the separation $\pm\p{p-q}$.  In this case, all $h_p=h_0$
 are equal, and \eqref{eq:cond_2_ref_start} can be written in the form
@@
 \end{align}
-We can reduce this eigenvalue problem even further by making use of
-reflection symmetries such as $h_{d,0}=h_{-d,0}$.  In $D$ dimensions,
-there is one such symmetry for each lattice axis, which implies that
-the degrees of freedom in $\ut{\v{\m w}}$ gets reduced by a factor of
-$\sim 2^D$.  Reflection symmetries similarly reduce the degrees of
+We can reduce this eigenvalue problem even further in the case of an
+isotropic system, in which $h_{pq}$ and $w_{pq}$ depend only on the
+distance $\abs{p-q}$.  On a $D$ dimensional lattice, isotropy implies
+one reflection symmetry for each lattice axis, which implies that the
+number of degrees of freedom in $\ut{\v{\m w}}$ gets reduced by a
+factor of $\sim 2^D$.  Isotropy similarly reduces the degrees of
 freedom in $\ut{\check{\m h}}$ from $\sim N\times N$ to

### 2020-01-18 collective_sun/multilevel_spin_notes.tex
fixed some errors left over from a change in notation
fixed minor mistake
removed derivations of effective Hamiltonians that are redundant with the work in other notes minor changes to notation, minor typo fixes

@@
 ways to assign $N$ subsystems to $n$ states, which is
-${N+n-1 \choose n-1}$, and the energy of any fully symmetric state
+${N+n-1 \choose n-1}$.  The energy of any fully symmetric state
 $\ket{\psi_\FS}\in\M_\FS$ with respect to the interaction Hamiltonian
 $H_{\t{int}}$ in \eqref{eq:H_int_perm} is
-$-u{N \choose 2}=-\p{1/4}N\p{N-1}u$.
+$-u{N \choose 2}=-\p{1/2}N\p{N-1}u$.
 
@@
   B_L^{(\phi,q)} = -t w_L\p{qa} A_L^{(\phi)},
+  \label{eq:B_L_phi}
 \end{align}
@@
   H_{\t{Raman,bare}}^{(\phi)}
-  = \sum_j \sum_{m\in\set{1,0,-1}} \p{\Omega_m e^{i\phi_m j}
+  = \sum_j \sum_{m\in\set{1,0,-1}} \Omega_m \p{e^{i\phi_m j}
     T_{1,m}^{(j)} \otimes \op{\up}{\dn} + \t{h.c.}}
@@
 \end{align}
-where $\omega$ is a driving frequency, and $\chi_p$ is a (real)
-dimensionless driving amplitude for spin $p$, and
+where $\omega$ is a driving frequency, $\chi_p$ is a dimensionless
+driving amplitude for spin $p$, and
 $D_\z^{(p)} \equiv \xi_1 D_{1,0}^{(p)}$ with $\xi_1$ defined in
@@
 \end{align}
-
-\vspace{5cm}
 
@@
   \begin{itemize}
-  \item Special angles $\phi$
+  \item Special angles $\phi$?
   \end{itemize}
+\item Different (more realistic / general) drives
 \item Generalizations of squeezing beyond SU(2)?
@@
   \item Super-exchange regime
-  \item Non-uniform (``disordered'') interactions
+  \item Non-uniform (local / disordered) interactions
   \end{itemize}
-\item Consider different drives
 \end{itemize}
@@
   T_{\ell_1 m_1} T_{\ell_2 m_2}
-  = \sum_{L,M} f_{\ell_1 m_1;\ell_2 m_2}^{LM} T_{LM},
+  = \sum_{L,M} g_{\ell_1 m_1;\ell_2 m_2}^{LM} T_{LM},
 \end{align}
@@
 \begin{align}
-  f_{\ell_1 m_1;\ell_2 m_2}^{LM}
+  g_{\ell_1 m_1;\ell_2 m_2}^{LM}
   \equiv \obk{T_{LM} | T_{\ell_1 m_1} T_{\ell_2 m_2}}
@@
 \begin{multline}
-  f_{\ell_1 m_1;\ell_2 m_2}^{LM}
+  g_{\ell_1 m_1;\ell_2 m_2}^{LM}
   = \delta_{M,m_1+m_2}
@@
 \begin{align}
-  f_{\ell_1 m_1;\ell_2 m_2}^{LM}
+  g_{\ell_1 m_1;\ell_2 m_2}^{LM}
   &= \delta_{M,m_1+m_2} \p{-1}^{2I+M}
@@
 \begin{align}
-  f_{\ell_1 m_1;\ell_2 m_2}^{LM}
-  = \p{-1}^{\ell_1+\ell_2-L} f_{\ell_2 m_2;\ell_1 m_1}^{LM}.
+  g_{\ell_1 m_1;\ell_2 m_2}^{LM}
+  = \p{-1}^{\ell_1+\ell_2-L} g_{\ell_2 m_2;\ell_1 m_1}^{LM}.
 \end{align}
@@
   = \f{\tau_{m_1}\tau_{m_2}}{\sqrt{2}}
-  \sum_L \p{f_{\ell_1 m_1;\ell_2 m_2}^{L,m_1+m_2}
+  \sum_L \p{g_{\ell_1 m_1;\ell_2 m_2}^{L,m_1+m_2}
     E_{\ell_1 m_1;\ell_2 m_2}^{(L,+)}
-    + f_{\ell_1 m_1;\ell_2,-m_2}^{L,m_1-m_2}
+    + g_{\ell_1 m_1;\ell_2,-m_2}^{L,m_1-m_2}
     E_{\ell_1 m_1;\ell_2 m_2}^{(L,-)}},
@@
   D_{\ell_1,0} D_{\ell_2 m}
-  = \sum_L f_{\ell_1,0,\ell_2,m}^{Lm}
+  = \sum_L g_{\ell_1,0,\ell_2,m}^{Lm}
   \sum_{s\in\set{\pm1}} \f12 \sp{1 + s \p{-1}^{L+\ell_1+\ell_2}}
@@
 
-Here we derive the effective Hamiltonian induced on the fully
-symmetric manifold $\M_\FS$ by the single-particle SOC Hamiltonian
+Here we derive the first- and second-order effective Hamiltonians
+induced on the fully symmetric manifold $\M_\FS$ by the
+single-particle weak-SOC Hamiltonian
 \begin{align}
   H_{\t{lat}}^{(\phi)}
-  = -t \sum_{q,L} w_L\p{q} A_L^{(\phi)} D_{L,0}^{(q)},
-  &&
-  w_L\p{q} \equiv
-  \begin{cases}
-    \cos q & L~\t{even} \\
-    \sin q & L~\t{odd}
-  \end{cases},
-\end{align}
-where $A_L^{(\phi)}$ is essentially a sum over Clebsch-Gordan
-coefficients, defined in \eqref{eq:A_L_phi}, and scales as
-$A_L^{(\phi)} \sim \phi^L + O\p{\phi^{L+2}}$ when $\abs{\phi}\ll1$.
-For brevity, we work in units with the lattice spacing $a=1$ and
-implicitly assume the limit $\abs{\phi}\ll1$ throughout this section.
-
-In order to simplify the calculation of effective Hamiltonians, we
-write the SOC Hamiltonian in the form
-\begin{align}
-  H_{\t{lat}}^{(\phi)} = \sum_{q,L} v_L^{(q)} D_{L,0}^{(q)},
-  &&
-  v_L^{(q)} \equiv - t w_L\p{q} A_L^{(\phi)},
-\end{align}
-and decompose it into its uniform and inhomogeneous parts as
-\begin{align}
-  H_{\t{lat}}^{(\phi)}
-  = \sum_L v_L \D_{L,0} + \sum_{q,L} z_L^{(q)} D_{L,0}^{(q)},
-\end{align}
-where
-\begin{align}
-  v_L \equiv \EE_q\sp{v_L^{(q)}},
-  &&
-  z_L^{(q)} \equiv v_L^{(q)} - v_L,
-  &&
-  \EE_q\sp{X_q} \equiv \f1N \sum_q X_q.
-\end{align}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Effective SOC at first order}
-\label{sec:H_eff_1}
-
-The effective Hamiltonian in induced on the fully symmetric manifold
-$\M_\FS$ at first order is
+  = \sum_{q,L} B_L^{(\phi,q)} D_{L,0}^{(q)},
+\end{align}
+where $D_{L,0}^{(q)}$ is a drive operator that acts on spin $q$, and
+$B_L^{(\phi,q)}$ is an effective inhomogeneous driving field amplitude
+defined in \eqref{eq:B_L_phi} and \eqref{eq:A_L_phi}.  A perturbative
+treatment of $H_{\t{lat}}^{(\phi)}$ yields the effective
+Hamiltonians[\red{TODO: cite perturbation theory notes}]
 \begin{align}
   H_{\t{eff}}^{(1)}
-  = \P_0 H_{\t{lat}}^{(\phi)} \P_0
-  = \sum_L v_L \P_0 \D_{L,0} \P_0
-  + \sum_{q,L} z_L^{(q)} \P_0 D_{L,0}^{(q)} \P_0,
-\end{align}
-where $\P_0$ is a projector onto $\M_\FS$.  Permutational symmetry
-within the fully symmetric manifold $\M_\FS$ allow us to freely change
-the quasi-momentum index on $\P_0D_{L,0}^{(q)}\P_0$, and the
-coefficients $z_L^{(q)}$ are mean-zero by construction, so
-\begin{align}
-  \sum_{q,L} z_L^{(q)} \P_0 D_{L,0}^{(q)} \P_0
-  = \sum_q z_L^{(q)} \P_0 D_{L,0}^{(0)} \P_0
-  = 0,
-\end{align}
-which implies that
-\begin{align}
-  H_{\t{eff}}^{(1)}
-  = \sum_L v_L \D_{L,0}
-  = -t \sum_L \EE_q\sp{w_L\p{q}} A_L^{(\phi)} \D_{L,0},
-  \label{eq:H_eff_1}
-\end{align}
-where we implicitly restrict $H_{\t{eff}}^{(1)}$ to the fully
-symmetric manifold $\M_\FS$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Effective SOC at second order}
-\label{sec:H_eff_2}
-
-The effective Hamiltonian in induced on the fully symmetric manifold
-$\M_\FS$ at second order is\cite{bravyi2011schrieffer,
-  perlin2019effective}
-\begin{align}
+  &\equiv \sum_L \EE_q\sp{B_L^{(\phi,q)}} \D_{L,0},
+  \\
   H_{\t{eff}}^{(2)}
-  = - \P_0 H_{\t{lat}}^{(\phi)} \E H_{\t{lat}}^{(\phi)} \P_0
-  &&
-  \E \equiv \sum_{\Delta>0} \f{\P_\Delta}{\Delta},
-\end{align}
-where $\P_\Delta$ is a projector onto the eigenspace of the
-interaction Hamiltonian $H_0$, with interaction energy $\Delta$ above
-that of fully symmetric manifold $\M_\FS$.  The uniform part of the
-SOC Hamiltonian $H_{\t{lat}}^{(\phi)}$ is permutationally symmetric,
-and therefore does not couple the fully symmetric manifold $\M_\FS$ to
-its orthogonal complement.  As a consequence, the uniform part of
-$H_{\t{lat}}^{(\phi)}$ is annihilated between the projector $\P_0$
-onto $\M_\FS$ and the projectors in $\E$ onto the orthogonal
-complement of $\M_\FS$.  We can therefore keep only the inhomogeneous
-parts of $H_{\t{lat}}^{(\phi)}$ and expand
-\begin{align}
-  H_{\t{eff}}^{(2)}
-  = -\sum_{J,K} \P_0 \Z_J \E \Z_K \P_0,
-  &&
-  \Z_L \equiv \sum_{q,L} z_L^{(q)} D_{L,0},
-  \label{eq:H_eff_2_start}
-\end{align}
-We now divide the task of further simplifying $H_{\t{eff}}^{(2)}$ into
-two steps:
-\begin{enumerate*}
-\item eliminating the excited-state projectors in $\E$, and
-\item projecting the remaining product of operators in
-  \eqref{eq:H_eff_2_start} onto the fully symmetric manifold.
-\end{enumerate*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Eliminating excited-state projectors}
-
-In order to eliminate the excited-state projectors in $\E$, we observe
-that the action of $\Z_L$ on any fully symmetric state
-$\ket{\psi_{\FS}}\in\M_\FS$ yields an eigenvector of the interaction
-Hamiltonian $H_{\t{int}}=-u\sum_{p<q}\Pi_{pq}$.  To substantiate this
-claim, we expand
-\begin{align}
-  \sum_{p<q} \Pi_{pq} \Z_L \ket{\psi_\FS}
-  = \sum_{\substack{p<q\\k}}
-  z_L^{(k)} \Pi_{pq} D_{L,0}^{(k)} \ket{\psi_\FS}
-\end{align}
-If $k\notin\set{p,q}$, then the permutation operator $\Pi_{pq}$
-commutes with $D_{L,0}^{(k)}$, and in turn acts trivially on the
-permutationally-symmetric state $\ket{\psi_\FS}$, giving us
-\begin{align}
-  \sum_{\substack{p<q\\k\notin\set{p,q}}}
-  z_L^{(k)} \Pi_{pq} D_{L,0}^{(k)} \ket{\psi_\FS}
-  = \sum_k \sum_{\substack{p<q\\p,q\ne k}}
-  z_L^{(k)} D_{L,0}^{(k)} \ket{\psi_\FS}
-  = { N-1 \choose 2 } \Z_L \ket{\psi_\FS}.
-  \label{eq:Z_eig_notin}
-\end{align}
-If $k\in\set{p,q}$, then we can resolve the identity operator
-$\1=\Pi_{pq}\Pi_{pq}$ to expand
-\begin{multline}
-  \sum_{\substack{p<q\\k\in\set{p,q}}}
-  z_L^{(k)} \Pi_{pq} D_{L,0}^{(k)} \ket{\psi_\FS}
-  = \f12 \sum_{\substack{p\ne q\\k\in\set{p,q}}}
-  z_L^{(k)} \Pi_{pq} D_{L,0}^{(k)} \Pi_{pq} \Pi_{pq} \ket{\psi_\FS}
-  = \sum_q \sum_{k\ne q} z_L^{(k)} D_{L,0}^{(q)} \ket{\psi_\FS},
-\end{multline}
-where, using the fact that the coefficients $z_L^{(k)}$ are mean-zero,
-\begin{align}
-  \sum_{k\ne q} z_L^{(k)}
-  = \sum_k z_L^{(k)} - z_L^{(q)}
-  = - z_L^{(q)},
-\end{align}
-so
-\begin{align}
-  \sum_{\substack{p<q\\k\in\set{p,q}}}
-  z_L^{(k)} \Pi_{pq} D_{L,0}^{(k)} \ket{\psi_\FS}
-  = - \sum_q z_L^{(q)} D_{L,0}^{(q)} \ket{\psi_\FS}
-  = - \Z_L \ket{\psi_\FS}.
-\end{align}
-In total,
-\begin{align}
-  \sum_{p<q} \Pi_{pq} \Z_L \ket{\psi_\FS}
-  = \sp{{N-1\choose 2}-1} \Z_L \ket{\psi_\FS}.
-\end{align}
-The image of any fully symmetric state $\ket{\psi_\FS}\in\M_\FS$ under
-the action of $\Z_L$ is thus proportional to an eigenstate of
-$H_{\t{int}}$ with interaction energy $-u\sp{{ N-1 \choose 2 } - 1}$,
-which is $Nu$ above the interaction energy $-u{N\choose2}$ of states
-in the fully symmetric manifold, so
-\begin{align}
-  \E \Z_L \P_0
-  = \sum_{\Delta>0} \f1{\Delta} \P_\Delta \Z_L \P_0
-  =  \f1{Nu} \Z_L \P_0
-  = \f1{fU} \Z_L \P_0,
-  \label{eq:proj_elim}
-\end{align}
-where $f$ is the filling fraction of lattice sites and $U$ is the
-two-body on-site interaction energy.  Our effective Hamiltonian thus
-far is
-\begin{align}
-  H_{\t{eff}}^{(2)}
-  = - \f1{fU} \sum_{J,K} \P_0 \Z_J \Z_K \P_0.
-  \label{eq:H_eff_2_mid}
-\end{align}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Projecting onto the fully symmetric manifold}
-
-We now need to evaluate the product
-\begin{align}
-  \P_0 \Z_J \Z_K \P_0
-  = \sum_{p,q} z_J^{(p)} z_K^{(q)} \P_0 D_{J,0}^{(p)} D_{K,0}^{(q)} \P_0.
-  \label{eq:PZZP_start}
-\end{align}
-Using the permutational symmetry of the fully symmetric manifold, we
-can simplify the $p=q$ terms above to
-\begin{align}
-  \sum_q z_J^{(q)} z_K^{(q)}
-  \P_0 D_{J,0}^{(q)} D_{K,0}^{(q)} \P_0
-  = N \cov_q\sp{v_J^{(q)},v_K^{(q)}}
-  \P_0 D_{J,0}^{(0)} D_{K,0}^{(0)} \P_0,
-  \label{eq:PZZP_eq}
-\end{align}
-where
+  &\equiv \sum_{J,K}
+  \f{\cov_q\sp{B_J^{(\phi,q)}, B_K^{(\phi,q)}}}{u\p{N-1}}
+  \p{\D_{J,0} \D_{K,0} - N \sum_L g_{JKL} \D_{L,0}},
+\end{align}
+where $\EE_q\sp{X_q} \equiv \sum_q X_q / N$ is the mean of $X_q$ over
+all occupied quasi-momenta $q$;
 \begin{align}
@@
   \equiv \EE_q\sp{\p{X_q-\EE_k\sp{X_k}}\p{Y_q-\EE_\ell\sp{Y_\ell}}}
-  = \EE_q\sp{X_q Y_q} - \EE_k\sp{X_k} \EE_\ell\sp{Y_\ell}
+  = \EE_q\sp{X_q Y_q} - \EE_k\sp{X_k} \EE_\ell\sp{Y_\ell},
   \label{eq:cov}
 \end{align}
-is the covariance between $X_q$ and $Y_q$, evaluated over all occupied
-quasi-momenta $q$.  The terms with $p\ne q$ in \eqref{eq:PZZP_start},
-meanwhile, are
-\begin{align}
-  \sum_{p\ne q} z_J^{(p)} z_K^{(q)}
-  \P_0 D_{J,0}^{(p)} D_{K,0}^{(q)} \P_0
-  = \sum_{p\ne q} z_J^{(p)} z_K^{(q)}
-  \P_0 D_{J,0}^{(0)} D_{K,0}^{(1)} \P_0,
-\end{align}
-where the remaining sum over coefficients is
-\begin{align}
-  \sum_{p\ne q} z_J^{(p)} z_K^{(q)}
-  = \sum_{p,q} z_J^{(p)} z_K^{(q)} - \sum_q z_J^{(q)} z_K^{(q)}
-  = - N \cov_q\sp{v_J^{(q)},v_K^{(q)}},
-\end{align}
-so
-\begin{align}
-  \sum_{p\ne q} z_J^{(p)} z_K^{(q)}
-  \P_0 D_{J,0}^{(p)} D_{K,0}^{(q)} \P_0
-  = - N \cov_q\sp{v_J^{(q)},v_K^{(q)}}
-  \P_0 D_{J,0}^{(0)} D_{K,0}^{(1)} \P_0.
-  \label{eq:PZZP_neq}
-\end{align}
-Combining \eqref{eq:PZZP_eq} and \eqref{eq:PZZP_neq} gives us
-\begin{align}
-  \P_0 \Z_J \Z_K \P_0
-  = N \cov_q\sp{v_J^{(q)},v_K^{(q)}}
-  \P_0\p{D_{J,0}^{(0)} D_{K,0}^{(0)}
-    - D_{J,0}^{(0)} D_{K,0}^{(1)}} \P_0.
-\end{align}
-In order to express this result in terms of collective operators, we
-expand
-\begin{align}
-  \P_0 \D_{J,0} \D_{K,0} \P_0
-  = N \P_0 D_{J,0}^{(0)} D_{K,0}^{(0)} \P_0
-  + N\p{N-1} \P_0 D_{J,0}^{(0)} D_{K,0}^{(1)} \P_0,
-\end{align}
-which implies that
-\begin{align}
-  - N \P_0 D_{J,0}^{(0)} D_{K,0}^{(1)} \P_0
-  = \f1{N-1} \P_0 \p{N D_{J,0}^{(0)} D_{K,0}^{(0)}
-    - \D_{J,0} \D_{K,0}} \P_0
-\end{align}
-and in turn
-\begin{align}
-  \P_0 \Z_J \Z_K \P_0
-  = \f1{N-1} \cov_q\sp{v_J^{(q)},v_K^{(q)}}
-  \P_0 \p{N^2 D_{J,0}^{(0)} D_{K,0}^{(0)} - \D_{J,0} \D_{K,0}} \P_0.
-\end{align}
-Finally, we can simplify
-\begin{align}
-  N \P_0 D_{J,0}^{(0)} D_{K,0}^{(0)} \P_0
-  = N \sum_L f_{JKL} \P_0 D_{L,0}^{(0)} \P_0
-  = \sum_L f_{JKL} \P_0 \D_{L,0} \P_0,
-  &&
-  f_{JKL} \equiv f_{J,0;K,0}^{L,0},
-\end{align}
-where $f_{\ell_1 m_1;\ell_2 m_2}^{LM}$ is a structure constant of the
-transition operator algebra, defined in \eqref{eq:trans_prod}.  In
-total,
-\begin{align}
-  \P_0 \Z_J \Z_K \P_0
-  = \f1{N-1} \cov_q\sp{v_J^{(q)},v_K^{(q)}}
-  \P_0 \p{N \sum_L f_{JKL} \D_{L,0} - \D_{J,0} \D_{K,0}} \P_0,
-\end{align}
-which implies that the effective Hamiltonian $H_{\t{eff}}^{(2)}$ in
-\eqref{eq:H_eff_2_mid} is
-\begin{align}
+is the covariance of $X_q$ and $Y_q$ over all occupied quasi-momenta
+$q$; and $g_{JKL} \equiv g_{J,0;K,0}^{L,0}$ is a structure constant of
+the transition operator algebra, defined in \eqref{eq:trans_prod}.
+Substituting the form of $B_L^{(\phi,q)}$ in \eqref{eq:B_L_phi}, we
+get
+\begin{align}
+  H_{\t{eff}}^{(1)}
+  &= -t \sum_L \EE_q\sp{w_L\p{qa}} A_L^{(\phi)} \D_{L,0},
+  \label{eq:H_eff_1}
+  \\
   H_{\t{eff}}^{(2)}
-  = \f{t^2}{fU\p{N-1}} \sum_{J,K}
-  \cov_q\sp{w_J\p{qa},w_K\p{qa}} A_J^{(\phi)} A_K^{(\phi)}
-  \p{\D_{J,0} \D_{K,0} - N \sum_L f_{JKL} \D_{L,0}},
+  &\equiv \f{t^2}{fU\p{N-1}} \sum_{J,K}
+  \cov_q\sp{w_J\p{qa}, w_K\p{qa}} A_J^{(\phi)} A_K^{(\phi)}
+  \p{\D_{J,0} \D_{K,0} - N \sum_L g_{JKL} \D_{L,0}}.
   \label{eq:H_eff_2}
 \end{align}
-where we have recovered factors of the lattice spacing $a$ as
-appropriate, and we have dropped the explicit projection operators
-$\P_0$ with the understanding that we are implicitly restricting all
-effective dynamics to the fully symmetric manifold $\M_\FS$.  Note
-that as the spatial filling fraction $f\to1$, the sums over occupied
-quasi-momenta $q$ become sums over nearly all angles
-$q\in\ZZ_N\times2\pi/N$.  In this case, when both $J,K$ are even or
-odd (i.e.~$J=K~\t{mod}~2$) the covariance in \eqref{eq:H_eff_2}
-becomes $\cov_p\sp{\sin\p{p},\sin\p{p}}\to1/2$ or
+where $t$ is the nearest-neighbor tunneling rate, $a$ is the spacing
+between neighboring lattice sites, $f$ is the filling fraction of
+spatial modes, $U$ is the two-body on-site interaction energy, and
+$w_L$ is cosine (sine) for even (odd) $L$.  Note that as the spatial
+filling fraction $f\to1$, the sums over occupied quasi-momenta $q$
+become sums over nearly all angles $q\in\ZZ_N\times2\pi/N$.  In this
+case, when both $J,K$ are even or odd (i.e.~$J=K~\t{mod}~2$) the
+covariance in \eqref{eq:H_eff_2} becomes
+$\cov_p\sp{\sin\p{p},\sin\p{p}}\to1/2$ or
 $\cov_p\sp{\cos\p{p},\cos\p{p}}\to1/2$, whereas when one of $J,K$ is
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Extracting dependence on the SOC angle $\phi$}
-
 The effective Hamiltonians in \eqref{eq:H_eff_1} and
 \eqref{eq:H_eff_2} are respectively first and second order in the SOC
-Hamiltonian $H_{\t{lat}}^{(\phi)}$.  Here we extract the leading and
+Hamiltonian $H_{\t{lat}}^{(\phi)}$.  We can extract the leading and
 next-to-leading order dependence of these effective Hamiltonians on
-the SOC angle $\phi$ using the expansion
+the SOC angle $\phi$ by using the expansions of $A_L^{(\phi)}$ in
+\eqref{eq:A_L_phi_small}, which gives us
 \begin{align}
@@
 \end{document}
+
+% Level statistics of Hamiltonian
+% Dynamics of single-particle observables
+% Dynamics of two-point correlators
+% - Different sites
+% - Two-time correlators
+% TWA dynamics

### 2020-01-21 soc_squeezing/collective_paper/review/response.tex
committing final changes to collective spin paper and codes

@@
       will rapidly take unphysical values with
-      $\abs{\bk{\hat\S_{\v n}\p{t}}}\gtrsim S^{\abs{\v n}}$.}  The
-    sudden and drastic departure from virtually exact results is
-    consistent with the limitations of the TST expansion discussed at
-    the end of Section II.  Specifically, we identify three possible
-    mechanisms for breakdown:
+      $\abs{\bk{\hat\S_{\v n}\p{t}}}\gtrsim S^{\abs{\v n}}$ (see
+      Appendix L for an example).}  The sudden and drastic departure
+    from virtually exact results is consistent with the limitations of
+    the TST expansion discussed at the end of Section II.
+    Specifically, we identify three possible mechanisms for breakdown:
     \begin{enumerate*}
@@
 
-  Note that we used of bold text to help identify the changes above;
-  the manuscript itself does not make these changes bold.  We hope
-  that these changes clarify the relationship between breakdown of the
-  TST expansion and an unphysical squeezing parameter.
+  Note that we used bold text to help identify the changes above; the
+  manuscript itself does not make these changes bold.  We have also
+  added a short new appendix (L) to provide an example of breakdown
+  without an unphysical squeezing parameter.  We hope that these
+  changes clarify subtleties related to the referee's question.
 
@@
 
-\item Concerning the final question, about the advantages of the TNT
-  model in the presence of decoherence:
+\item Concerning the final question, about squeezing with strong
+  decoherence:
 
@@
   than OAT or TAT in the presence of decoherence.  We have therefore
-  rewritten most of the last paragraph of section III, changing:
+  rewritten some of the last paragraph of section III (end of page 5,
+  beginning of page 6), replacing:
 
@@
     presented in Figure 1.  Strong-decoherence computations of this
-    sort ...}
+    sort were used to put lower bounds...}
 
-  to:
+  by:
 
@@
     more squeezing than the OAT or TAT models in the presence of
-    strong decoherence.  The better performance of TNT is a
+    strong decoherence.  The better performance of TNT is in part a
     consequence of the fact that TNT initially generates squeezing at
@@
     corroborate this finding with quantum trajectory simulations of a
-    small system in Appendix L.  Strong-decoherence computations of
-    the sort used for Figure 2 ...}
+    small system in Appendix M.  Strong-decoherence computations of
+    the sort used for Figure 2 put lower bounds...}
 
@@
   performance of TNT.  We hope that these additions address the
-  referee's questions and concerns about figure 2 to a satisfactory
-  degree.
-
-  % One consequence of applying our method in the large system, strong
-  % decoherence regime in figure 2 is that there is no way (barring a
-  % detailed analysis of error and breakdown in the TST expansion) to
-  % completely rule out numerical errors, as there are no other
-  % simulation methods available to corroborate the results.  We had
-  % hoped that the benchmarking in figure 1 and the discussion of
-  % breakdown mechanisms would assure readers that the TST expansion
-  % self-diagnoses its sudden departure from virtually exact results,
-  % thereby increasing confidence that the TST expansion can be
-  % trusted when no other methods are available.
+  referee's questions about figure 2 to a satisfactory degree.
 

### 2020-01-21 soc_squeezing/collective_paper/collective_spin.tex
committing final changes to collective spin paper and codes

@@
 The Hamiltonians in Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT} squeeze the
-initial product state $\ket{\X}\propto\p{\ket\up+\ket\dn}^{\otimes N}$
-with $\hat S_\x\ket{\X}=S\ket{\X}$.  Our measure of spin squeezing is
-the directionally-unbiased Ramsey squeezing parameter determined by
-the maximal gain in resolution $\Delta\phi$ of a phase angle $\phi$
-over that achieved by any spin-polarized product state
-(e.g.~$\ket{\X}$)\cite{wineland1992spin, ma2011quantum},
+initial product state $\ket\X\propto\p{\ket\up+\ket\dn}^{\otimes N}$
+with $\hat S_\x\ket\X=S\ket\X$.  Our measure of spin squeezing is the
+directionally-unbiased Ramsey squeezing parameter determined by the
+maximal gain in resolution $\Delta\phi$ of a phase angle $\phi$ over
+that achieved by any spin-polarized product state
+(e.g.~$\ket\X$)\cite{wineland1992spin, ma2011quantum},
 \begin{align}
@@
   $\gamma_-=\gamma_+=\gamma_\z=\chi$
-  \label{fig:squeezing_incoherent}]
+  \label{fig:squeezing_dec}]
   {\includegraphics{decoherence_weak.pdf}}
@@
     under \sref{fig:squeezing_unitary} unitary and
-    \sref{fig:squeezing_incoherent} non-unitary dynamics, computed
-    using exact methods (solid lines), quantum trajectory simulations
+    \sref{fig:squeezing_dec} non-unitary dynamics, computed using
+    exact methods (solid lines), quantum trajectory simulations
     (dots), and the TST expansion in Eq.~\eqref{eq:TST} with $M=35$
@@
 Figure \ref{fig:benchmarking} compares the squeezing parameter $\xi^2$
-for $N=10^4$ spins initially in the state $\ket{\X}$ evolved under the
+for $N=10^4$ spins initially in the state $\ket\X$ evolved under the
 Hamiltonians in Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT}, as computed via
@@
 dephasing of individual spins at rates $\chi$ (Figure
-\ref{fig:squeezing_incoherent}), respectively described by the sets of
-jump operators $\J_\a\equiv\set{\hat s_\a^{(j)}}$ with corresponding
+\ref{fig:squeezing_dec}), respectively described by the sets of jump
+operators $\J_\a\equiv\set{\hat s_\a^{(j)}}$ with corresponding
 decoherence rates $\gamma_\a=\chi$ for $\a\in\set{-,+,\z}$.  The
@@
 rapidly take unphysical values with
-$\abs{\bk{\hat\S_{\v n}\p{t}}}\gtrsim S^{\abs{\v n}}$.  The sudden and
-drastic departure from virtually exact results is consistent with the
-limitations of the TST expansion discussed at the end of Section
-\ref{sec:theory}.  Specifically, we identify three possible mechanisms
-for breakdown:
+$\abs{\bk{\hat\S_{\v n}\p{t}}}\gtrsim S^{\abs{\v n}}$ (see Appendix
+\ref{sec:breakdown} for an example).  The sudden and drastic departure
+from virtually exact results is consistent with the limitations of the
+TST expansion discussed at the end of Section \ref{sec:theory}.
+Specifically, we identify three possible mechanisms for breakdown:
 \begin{enumerate*}
@@
 fewer computing resources.  The TST expansion results in Figure
-\ref{fig:squeezing_incoherent}, for example, take $\sim10$ seconds to
-compute with a single CPU on modern computing hardware.  The quantum
+\ref{fig:squeezing_dec}, for example, take $\sim10$ seconds to compute
+with a single CPU on modern computing hardware.  The quantum
 trajectory Monte Carlo results in the same figure, meanwhile, take
@@
 \ref{fig:decoherence_strong} shows squeezing of $N=10^4$ spins
-initially in $\ket{\X}$, undergoing spontaneous decay, excitation, and
+initially in $\ket\X$, undergoing spontaneous decay, excitation, and
 dephasing of individual spins at rates
@@
 model can generate more squeezing than the OAT or TAT models in the
-presence of strong decoherence.  The better performance of TNT is a
-consequence of the fact that TNT initially generates squeezing at a
-faster rate than OAT or TAT, thereby allowing it to produce more
+presence of strong decoherence.  The better performance of TNT is in
+part a consequence of the fact that TNT initially generates squeezing
+at a faster rate than OAT or TAT, thereby allowing it to produce more
 squeezing before the degrading effects of decoherence kick in.  We
 corroborate this finding with quantum trajectory simulations of a
-small system in Appendix \ref{sec:small_system}.  Strong-decoherence
+smaller system in Appendix \ref{sec:strong_dec}.  Strong-decoherence
 computations of the sort used for Figure \ref{fig:decoherence_strong}
-were used to put lower bounds on theoretically achievable spin
-squeezing via TAT with decoherence in
-Ref.~[\citenum{he2019engineering}], exemplifying a concrete and
-practical application of the TST expansion and the collective-spin
-structure constants calculated in this work.
+put lower bounds on theoretically achievable spin squeezing via TAT
+with decoherence in Ref.~[\citenum{he2019engineering}], exemplifying a
+concrete and practical application of the TST expansion and the
+collective-spin structure constants calculated in this work.
 
@@
     spins initially in the polarized state
-    $\ket{\X}\propto\p{\ket\up+\ket\dn}^{\otimes N}$ evolving under
-    the squeezing Hamiltonians in Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT}.
+    $\ket\X\propto\p{\ket\up+\ket\dn}^{\otimes N}$ evolving under the
+    squeezing Hamiltonians in Eqs.~\eqref{eq:OAT}--\eqref{eq:TNT}.
     Results are shown for both unitary dynamics (solid lines) and
-    non-unitary dynamics with $\gamma_-=\gamma_+=\gamma_\z=100\chi$
-    (dashed lines), computed using the TST expansion in
-    Eq.~\eqref{eq:TST} with $M=20$.}
+    non-unitary dynamics with
+    $\gamma_-=\gamma_+=\gamma_\z=\gamma_0=100\chi$ (dashed lines),
+    computed using the TST expansion in Eq.~\eqref{eq:TST} with
+    $M=20$.}
   \label{fig:multi_time}
@@
 
-\vfill
 \onecolumngrid
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Robustness of squeezing generated by the twist-and-turn (TNT)
-  model}
-\label{sec:small_system}
-
-% \begin{figure}
-%   \centering
-%   \includegraphics{decoherence_strong_small.pdf}
-%   \caption{Spin squeezing of $N=100$ spins initially in $\ket\X$ with
-%     spontaneous decay, excitation, and dephasing of individual spins
-%     at rates $\gamma_-=\gamma_+=\gamma_\z=2\chi$.  Computed computed
-%     using exact methods (solid line), quantum trajectory simulations
-%     with $10^4$ trajectories (dots), and the TST expansion in
-%     Eq.~\eqref{eq:TST} with $M=35$ (dashed lines).}
-%   \label{fig:decoherence_strong_small}
-% \end{figure}
+\section{Diagnosing breakdown of the TST expansion}
+\label{sec:breakdown}
+
+In Figure \ref{fig:benchmarking} of the main text, the TST expansion
+provided nearly exact results for squeezing until a sudden departure
+that quickly resulted in an unphysical squeezing parameter, $\xi^2<0$.
+In general, however, there is no fundamental relationship between the
+breakdown of the TST expansion and the conditions for a physical
+squeezing parameter $\xi^2$.  A proper diagnosis of breakdown
+therefore requires inspection of the correlators $\bk{\S_{\v n}\p{t}}$
+used to compute the squeezing parameter $\xi^2$, which upon breakdown
+will rapidly take unphysical values with
+$\abs{\bk{\S_{\v n}\p{t}}}\gtrsim S^{\abs{\v n}}$.  As an example,
+Figure \ref{fig:unitary_small} shows the squeezing parameter $\xi^2$
+throughout decoherence-free evolution of $N=100$ spins initially in
+the state $\ket\X$.  In this example, the squeezing computed by the
+TST expansion for the TAT model diverges from the exact answer without
+an immediate and obvious signature of breakdown.  Nonetheless,
+breakdown can still be diagnosed by inspection of individual
+correlators, as shown in Figure \ref{fig:correlator_example}, where we
+plot $\t{Im}\bk{S_+^2}$ as a function of time for $N=100$ spins
+evolving under the TAT without decoherence.  Figure
+\ref{fig:correlator_example} shows that breakdown clearly occurs
+around $N\chi t\lesssim7$, when the correlator $\bk{S_+^2}$ begins to
+diverge to values $\gtrsim S^2=\p{N/2}^2=2500$ in magnitude.  A joint
+inspection of figures \ref{fig:unitary_small} and
+\ref{fig:correlator_example} suffice to trace the anomalous behavior
+of $\xi^2$ from $N\chi t\approx 7$ back to $N\chi t\approx 6$, when it
+first took a sudden turn before becoming unphysical at
+$N\chi t\approx 8$.
+
+\begin{figure}
+  \centering
+  \includegraphics{unitary_small.pdf}
+  \caption{Spin squeezing throughout decoherence-free evolution of
+    $N=100$ spins initially in the state $\ket\X$, computed using both
+    exact methods (solid lines) and the TST expansion in
+    Eq.~\eqref{eq:TST} with $M=35$ (dashed lines).  Solid circles mark
+    the times at which the TST expansion gives an unphysical result
+    with $\xi^2<0$.}
+  \label{fig:unitary_small}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics{correlator_example.pdf}
+  \caption{A collective spin correlator in the TAT model with $N=100$
+    spins and no decoherence, computed using the TST expansion with
+    $M=35$.  The divergence of correlators of this sort can be used to
+    diagnose the breakdown of the TST expansion.}
+  \label{fig:correlator_example}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Spin squeezing with strong decoherence}
+\label{sec:strong_dec}
+
+Here we provide supplementary evidence of our finding in Section
+\ref{sec:squeezing} that the TNT model can produce more squeezing than
+the OAT or TAT models in the presence of strong decoherence.  To this
+end, Figure \ref{fig:sqz_dec_scaling} shows the minimal squeezing
+parameter $\xi^2_{\t{min}}$ achievable with $N=100$ spins through the
+OAT, TAT, and TNT models as a function of the rate $\gamma_0$ at which
+individual spins undergo spontaneous decay, excitation, and dephasing.
+These results were computed with quantum trajectory simulations, with
+$10^3$ trajectories per data point.  While the OAT and TAT models
+produce more squeezing than TNT model with weak decoherence, this
+squeezing falls off faster with an increasing decoherence rate
+$\gamma_0$.  The relative robustness of TNT is in part a consequence
+of the fact that TNT initially generates squeezing at a faster rate
+than OAT or TAT, thereby allowing it to produce more squeezing before
+the degrading effects of decoherence kick in.
+
+\begin{figure}
+  \centering
+  \includegraphics{sqz_dec_scaling.pdf}
+  \caption{Optimal spin squeezing of $N=100$ spins undergoing
+    spontaneous decay, excitation, and dephasing at rates
+    $\gamma_-=\gamma_+=\gamma_\z=\gamma_0$, computed using quantum
+    trajectory simulations with $10^3$ trajectories per data point.}
+  \label{fig:sqz_dec_scaling}
+\end{figure}
 

### 2020-01-21 soc_squeezing/scripts/combine_trajectories.py

@@ -5,10 +5,13 @@ import numpy as np
 
 from squeezing_methods import squeezing_from_correlators
 
-log10_N = 3
-method = "TNT"
+log10_N = 2
+method = "OAT"
+dec_idx = 0
 
-files = glob.glob(f"../data/squeezing/jumps/sqz_D_exact_logN{log10_N}_{method}_s???.txt")
+data_dir = "../data/squeezing/jumps/"
+
+files = glob.glob(data_dir + f"sqz_D_exact_logN{log10_N}_{method}_d{dec_idx:02d}_s???.txt")
 data = sum([ np.loadtxt(f, dtype = complex, delimiter = ",")
              for f in files ]) / len(files)
 

### 2020-01-21 soc_squeezing/scripts/plot_squeezing.py

@@ -51,7 +51,7 @@ dec_rates_strong = [ (100,)*3, (0,)*3 ]
 
 OAT, TAT, TNT = "OAT", "TAT", "TNT"
 methods = [ OAT, TAT, TNT ]
-colors = { OAT: "#4E79A7", TAT: "#F28E2B", TNT : "#E15759" }
+colors = { OAT : "#1E4977", TAT : "#E15759", TNT : "#FFAE4B" }
 
 # construct Hamiltonians in (z,x,y) format, assuming an initial state in +X
 h_vec = { OAT : { (2,0,0) : 1 },
@@ -164,8 +164,6 @@ def jump_args(hamiltonian):
 for method in methods:
     if method == OAT: continue
     sqz_path = data_dir + f"sqz_D_exact_logN{log10_N}_{method}.txt"
-    if method == TNT:
-        sqz_path = data_dir + f"sqz_D_exact_logN{log10_N}_{method}.txt"
 
     if not os.path.isfile(sqz_path) or recompute_exact:
         correlators = correlators_from_trajectories(*jump_args(h_vec[method]))
@@ -240,7 +238,7 @@ for time_sqz in [ time_sqz_C_exact, time_sqz_C_trunc,
 time_pad = 1/3 # fractional time to add past TAT squeezing minimum
 sqz_pad = 1/10
 trajectory_marker_size = 2
-trunc_width = 2
+trunc_width = 3
 
 def positive(vals):
     idx = np.argmax(vals < 0)
@@ -258,13 +256,6 @@ def ylims(time_sqz_exact, time_sqz_trunc):
     yscale = (ymax/ymin)**sqz_pad
     return ymin/yscale, ymax*yscale
 
-# darken a color provided in hexidecimal format
-def darken_color(color, darken_val = 0.08):
-    # convert hexidecimal to [0,1)^3 RGB format
-    color = tuple([ int(color[jj:jj+2], 16)/255 - darken_val
-                           for jj in (1, 3 ,5) ])
-    return tuple( min(max(cc-darken_val,0), 1) for cc in color )
-
 ### unitary evolution
 
 plt.figure(figsize = figsize)
@@ -276,11 +267,11 @@ for method in methods:
     positive_vals = positive(time_sqz_C_trunc[method][1])
     plt.semilogy(time_sqz_C_trunc[method][0][:positive_vals],
                  time_sqz_C_trunc[method][1][:positive_vals],
[…]

### 2020-01-21 soc_squeezing/scripts/run_trajectories.py

@@ -1,30 +1,26 @@
 #!/usr/bin/env python3
 
 import sys
-import numpy as np
 
-from dicke_methods import coherent_spin_state
-from correlator_methods import mat_zxy_to_pzm, vec_zxy_to_pzm
-from jump_methods import correlators_from_trajectories
-
-assert(len(sys.argv[1:]) in [ 2, 3 ])
+assert(len(sys.argv[1:]) in [ 3, 4 ])
 method = sys.argv[1]
-seed = int(sys.argv[2])
+dec_idx = int(sys.argv[2])
+seed = int(sys.argv[3])
 
 OAT, TAT, TNT = "OAT", "TAT", "TNT"
 assert(method in [ OAT, TAT, TNT ])
 
-log10_N = 3
+log10_N = 2
 N = 10**log10_N
 
 trajectories = 100
 max_tau = 2
-time_steps = 100
+time_steps = 1000
 default_savepoints = True
 
-log_dir = "./logs/"
-data_dir = "../data/squeezing/jumps/"
-job_name = f"sqz_D_exact_logN{log10_N}_{method}_s{seed:03d}"
+log_dir = f"./logs/"
+data_dir = f"../data/squeezing/jumps/"
+job_name = f"sqz_D_exact_logN{log10_N}_{method}_d{dec_idx:02d}_s{seed:03d}"
 
 log_text = f"""#!/bin/sh
 
@@ -40,15 +36,24 @@ log_text = f"""#!/bin/sh
 
 module load python3
 
-python3 {sys.argv[0]} {method} {seed}
+python3 {sys.argv[0]} {method} {dec_idx} {seed}
 """
 
-if len(sys.argv[1:]) == 3:
+if len(sys.argv[1:]) == 4:
     with open(log_dir + job_name + ".sh", "w") as f:
         f.write(log_text)
     exit()
 
-dec_rates = [ (10,10,10), (0,0,0) ]
+######################################################################
+
+import numpy as np
+
+from dicke_methods import coherent_spin_state
+from correlator_methods import mat_zxy_to_pzm, vec_zxy_to_pzm
+from jump_methods import correlators_from_trajectories
+
+dec_rates = np.logspace(-3,1,13)
+dec_rates = [ (dec_rates[dec_idx],)*3, (0,0,0) ]
 
 init_state = "-Z"
 h_vec = {}

### 2020-01-21 soc_squeezing/scripts/squeezing_methods.py

@@ -1,238 +1,1 @@
-#!/usr/bin/env python3
-
-# FILE CONTENTS: model-independent methods for computing spin squeezing
-
-import numpy as np
-import scipy.sparse as sparse
-import scipy.linalg as linalg
-import scipy.optimize as optimize
-
-from correlator_methods import correlators_OAT
-
-##########################################################################################
-# general spin squeezing methods, and all necessary "infrastructure"
-##########################################################################################
-
-# act with a time-evolution unitary from the left, right, or both sides
-def evolve_left(state, hamiltonian, time):
-    return sparse.linalg.expm_multiply(-1j * time * hamiltonian, state)
-def evolve_right(state, hamiltonian, time):
-    return evolve_left(state.conj().T, hamiltonian, time).conj().T
-def evolve_left_right(state, hamiltonian, time):
-    return evolve_right(evolve_left(state, hamiltonian, time),
-                        hamiltonian, time)
-
-# time evolution of a state with a sparse hamiltonian
-def evolve(state, hamiltonian, time):
-    assert(sparse.issparse(hamiltonian))
-    assert(state.shape[0] == hamiltonian.shape[0])
-
-    if state.shape != hamiltonian.shape: # state is a vector
-        return evolve_left(state, hamiltonian, time)
-    else: # state is a density operator
-        return evolve_left_right(state, hamiltonian, time)
-
-# expectation value of X with respect to a state
-def val(X, state):
-    if X.shape != state.shape: # we have a state vector
-        return state.conj().T.dot(X.dot(state)).sum()
-    else: # we have a density operator
-        if sparse.issparse(X):
-            return X.T.multiply(state).sum()
-        else:
-            return (X.T*state).sum()
-
-# expectation value of vector spin operator
-def spin_vec_vals(state, S_op_vec):
-    return np.array([ val(X,state) for X in S_op_vec ])
-
-# expectation value of matrix spin-spin operator
[…]

### 2020-01-23 collective_sun/sun_perturbation.tex
consider the case of an inhomogeneous two-body perturbation in the TI analytics

@@
 diagonalizing the ${N \choose 2}\times{N \choose 2}\sim N^2\times N^2$
-matrix $\check{\m h}$.  When the interaction Hamiltonian $H_0$ is
+matrix $\check{\m h}$.  When both he interaction Hamiltonian $H_0$
 translationally invariant, however, this eigenvalue problem can be
@@
 We solve the single-body eigenvalue problem analytically, and reduce
-the two-body eigenvalue problem to that of diagonalizing an
-$\p{N-1}\times\p{N-1}$ matrix.  Finally, for we discuss the further
-reduction of the two-body problem to that of diagonalizing a matrix
+the two-body eigenvalue problem to that of diagonalizing
+$\p{N-1}\times\p{N-1}$ matrices.  Finally, for we discuss the further
+reduction of the two-body problem to that of diagonalizing matrices
 with dimensions $\sim\p{N/2^D}\times\p{N/2^D}$ in the case of an
-isotropic system on a $D$-dimensional periodic lattice, and we work
-out the case of $D=1$ explicitly as an example.
+isotropic system on a $D$-dimensional periodic lattice, but leave this
+reduction to future work.
 
@@
   \p{2h_{pq}-h_p-h_q} w_{pq}
-  + \sum_{k\notin\set{p,q}} \p{h_{pk} w_{kq} + w_{pk} h_{kq}}
+  + \sum_k \p{h_{pk} w_{kq} + w_{pk} h_{kq}}
   = \Delta w_{pq},
@@
 where $h_p\equiv\sum_q h_{pq}$.  If the interaction Hamiltonian $H_0$
-and perturbation $\V_2\p{\m w}$ are translationally invariant, then
-the couplings $h_{pq}$ and coefficients $w_{pq}$ depend on $p$ and $q$
-only through the separation $\pm\p{p-q}$.  In this case, all $h_p=h_0$
-are equal, and \eqref{eq:cond_2_ref_start} can be written in the form
-\begin{align}
-  2 \m h\circ \m w + \m h\c\m w + \m w\c\m h
-  = \sp{\Delta + 2h_0} \m w,
+is translationally invariant, then the couplings $h_{pq}$ depend on
+$p$ and $q$ only through the separation $\pm\p{p-q}$.  In this case,
+all $h_p=h_0$ are equal, and \eqref{eq:cond_2_ref_start} can be
+written in the form
+\begin{align}
+  2 \m h\circ \m w + \P_{\t{off-diag}}\p{\m h\c\m w + \m w\c\m h}
+  = \p{\Delta + 2h_0} \m w,
   \label{eq:cond_2_ref}
@@
 $\m h\circ\m w$ denotes an element-wise (Kronecker) product of $\m h$
-and $\m w$, i.e.~with matrix elements $h_{pq}w_{pq}$.  In order to
-find solutions to \eqref{eq:cond_2_ref}, we we expand the matrices
-$\m h,\m w$ in a basis that respects translational symmetry.  To this
-end, for all lattice site displacements $d\in\tilde\ZZ_L^D$ we define
-the {\it shift matrices}
-\begin{align}
-  \m s_d \equiv \sum_p \op{p+d}{p},
+and $\m w$, and
+$\P_{\t{off-diag}}\p{\m m}\equiv \m m-\sum_p\op{p}\m m\op{p}$ denotes
+the projection of $\m m$ onto the space of strictly off-diagonal
+matrices.  In order to find solutions to \eqref{eq:cond_2_ref}, we we
+expand the matrices $\m h,\m w$ in a basis that the respects
+translational symmetry of $\m h$.  To this end, for all lattice
+displacements $d\in\tilde\ZZ_L^D$ and wavenumbers
+$k\in\tilde\ZZ_L^D\times2\pi/L$ we define the {\it cycle matrices}
+\begin{align}
+  \m m_{dk} \equiv \sum_p e^{ip\c k} \op{p+d}{p},
 \end{align}
 where the element-wise sums in $p+d$ are implicitly taken modulo $L$
-(i.e.~the length of the periodic lattice along each axis).  The shift
-matrices allow us to expand
+(i.e.~the length of the periodic lattice along each axis).  In terms of the cycle matrices, we define
+\begin{align}
+  \tilde w_{dk} \equiv \f1N \tr\p{\m m_{dk}^\dag \m w},
+\end{align}
+expand
 \begin{align}
   \m h = \sum_{p,q} h_{pq} \op{p}{q}
-  = \sum_{d,q} h_{q+d,q} \op{q+d}{q}
-  = \sum_d h_{d,0} \m s_d,
-  &&
-  \m w = \sum_d w_{d,0} \m s_d,
-\end{align}
-and write the condition in \eqref{eq:cond_2_ref} as
-\begin{align}
-  \sum_{d\ne 0} h_{d,0} w_{d,0} \m s_d
-  + \sum_{\substack{c,d\ne0\\c+d\ne0}} h_{c,0} w_{d,0} \m s_{c+d}
-  = \p{\f12\Delta + h_0} \sum_{d\ne 0} w_{d,0} \m s_d,
-\end{align}
-where we explicitly exclude terms that always vanish due to the fact
-that $\m h,\m w$ are zero on the diagonal.  With some re-indexing, we
-can write
-\begin{align}
-  \sum_{d\ne 0} h_{d,0} w_{d,0} \m s_d
-  + \sum_{\substack{c,d\ne0}} h_{d-c,0} w_{c,0} \m s_d
-  = \p{\f12\Delta + h_0} \sum_{d\ne 0} w_{d,0} \m s_d,
-  \label{eq:cond_2_shift}
+  = \sum_{q,d} h_{q+d,q} \op{q+d}{q}
+  = \sum_d h_{d,0} \m m_{d,0},
+  &&
+  \m w = \sum_{d,k} \tilde w_{dk} \m m_{dk},
+\end{align}
+and write the condition in \eqref{eq:cond_2_ref} as a new set of
+conditions
+\begin{align}
+  \sum_{d\ne0} 2 h_{d,0} \tilde w_{dk} \m m_{dk}
+  + \sum_{\substack{c,d\\c+d\ne0}}
+  h_{c,0} \tilde w_{dk} \p{1 + e^{-ic\c k}} \m m_{c+d,k}
+  = \p{\Delta + 2 h_0} \sum_{d\ne0} \tilde w_{dk} \m m_{dk},
+\end{align}
+that must hold for all wavenumbers $k$.  With some re-indexing, we can
+write
+\begin{align}
+  \sum_{d\ne0} 2 h_{d,0} \tilde w_{dk} \m m_{dk}
+  + \sum_{c,d\ne0} h_{d-c,0} \tilde w_{ck}
+  \p{1 + e^{-i\p{d-c}\c k}} \m m_{dk}
+  = \p{\Delta + 2 h_0} \sum_{d\ne0} \tilde w_{dk} \m m_{dk},
+  \label{eq:cond_2_cycle}
 \end{align}
@@
 \begin{align}
-  \ut{\check{\m h}}
-  \equiv \sum_{d\ne0} h_{d,0} \op{d}
-  + \sum_{c,d\ne0} h_{d-c,0} \op{d}{c},
-  &&
-  \ut{\v{\m w}} \equiv \sum_{d\ne0} w_{d,0} \ket{d},
-\end{align}
-in order to express \eqref{eq:cond_2_shift} as the eigenvalue
-equation
-\begin{align}
-  \ut{\check{\m h}} \c \ut{\v{\m w}}
-  = \p{\f12\Delta + h_0} \ut{\v{\m w}}.
+  \ut{\check{\m h}}_k
+  \equiv \sum_{d\ne0} 2 h_{d,0} \op{d}
+  + \sum_{c,d\ne0} h_{d-c,0} \p{1+e^{-i\p{d-c}\c k}} \op{d}{c},
+  &&
+  \ut{\v{\m w}}_k \equiv \sum_{d\ne0} \tilde w_{dk} \ket{d},
+\end{align}
+in order to express \eqref{eq:cond_2_cycle} as the set of eigenvalue
+equations
+\begin{align}
+  \ut{\check{\m h}}_k \c \ut{\v{\m w}}_k
+  = \p{\Delta + 2 h_0} \ut{\v{\m w}}_k.
   \label{eq:cond_2_square}
 \end{align}
-We can reduce this eigenvalue problem even further in the case of an
-isotropic system, in which $h_{pq}$ and $w_{pq}$ depend only on the
-distance $\abs{p-q}$.  On a $D$ dimensional lattice, isotropy implies
-one reflection symmetry for each lattice axis, which implies that the
-number of degrees of freedom in $\ut{\v{\m w}}$ gets reduced by a
-factor of $\sim 2^D$.  Isotropy similarly reduces the degrees of
-freedom in $\ut{\check{\m h}}$ from $\sim N\times N$ to
-$\sim\p{N/2^D}\times\p{N/2^D}$.  We work out the case of $D=1$ below
-as an example, and leave the general case of arbitrary dimension
-$D\ge1$ to future work.
-
-In order to exploit reflection symmetry in one dimension ($D=1$), we
-define a complete basis of orthogonal vectors that are symmetric under
-reflection: for all $d\in\VV_N\equiv\set{1,2,\cdots,\floor{N/2}}$,
-\begin{align}
-  \ket{d_+} \equiv \f{\eta_d}{\sqrt{2}}\p{\ket{d} + \ket{-d}},
-  &&
-  \eta_d \equiv
-  \begin{cases}
-    1 & d \ne N/2, \\
-    1/\sqrt{2} & d = N/2
-  \end{cases},
-\end{align}
-where the special case of $d=N/2$ affects only even $N$, for which
-$\ket{N/2}=\ket{-N/2}$.  We can use these vectors to define a
-projector onto the reflection-symmetric subspace and expand
-$\ut{\v{\m w}}$ as
-\begin{align}
-  \P_+ \equiv \sum_{d\in\VV_N} \op{d_+},
-  &&
-  \ut{\v{\m w}} = \sum_{d\in\VV_N} w_{d,0} \,
-  \eta_d \sqrt{2} \ket{d_+}.
-\end{align}
-Finally, we project $\ut{\check{\m h}}$ onto the reflection-symmetric
-subspace,
-\begin{align}
-  \P_+ \ut{\check{\m h}} \P_+
-  &= \sum_{a,b\in\VV_N} \sum_{c,d\ne0}
-  \p{h_{d,0} \delta_{cd} + h_{c-d,0}} \op{a_+} \op{d} \op{b_+} \\
-  &= \sum_{a,b\in\VV_N} \op{a_+}{b_+} \f12 \eta_a\eta_b
-  \sum_{c,d\ne0} \p{h_{d,0} \delta_{cd} + h_{c-d,0}}
-  \p{\delta_{ac} + \delta_{-a,c}} \p{\delta_{bd} + \delta_{-b,d}} \\
-  &= \sum_{a\in\VV_N} h_{a,0} \op{a_+}
-  + \sum_{a,b\in\VV_N} \eta_a\eta_b
-  \p{h_{a-b,0} + h_{a+b,0}} \op{a_+}{b_+},
-\end{align}
-which allows us to write down the reduced eigenvalue problem
-\begin{align}
-  \P_+ \ut{\check{\m h}} \P_+ \c \ut{\v{\m w}}
-  = \p{\f12\Delta + h_0} \ut{\v{\m w}}.
-\end{align}
-Note that the factors of $\eta_a,\eta_b$ in
-$\P_+ \ut{\check{\m h}} \P_+$ only matter when $N$ is even, in which
-case these factors merely multiply the last row and last column of
-$\P_+ \ut{\check{\m h}} \P_+$ by a factor of $1/\sqrt{2}$.
+The appearance of a wavenumber index $k$ in \eqref{eq:cond_2_square}
+implies that all fixed-$k$ components of a general coefficient matrix
+$\m w$ must simultaneously satisfy \eqref{eq:cond_2_square} in order
+to generate states of fixed excitation energy $\Delta$ when the
+corresponding perturbation $\V_2^\Delta$ acts on a fully symmetric
+state $\ket\psi\in\M_0$.  The decomposition of an arbitrary
+coefficient vector $\v{\m w}=\sum_\Delta\v{\m w}_\Delta$ into
+coefficient vectors $\v{\m w}_\Delta$ that generate a fixed-energy
+perturbation $\V_2^\Delta$ is then performed by projecting each
+$k$-component of $\v{\m w}$ onto the corresponding $\Delta$-eigenspace
+of $\ut{\check{\m h}}_k$.  If the perturbation $\V_2$ is
+translationally invariant, then the $k$-components of $\v{\m w}$ for
+all $k\ne 0$ all vanish, so we only need to consider the case of
+$k=0$.
+
+We can reduce the eigenvalue problem in \eqref{eq:cond_2_square} even
+further in the case of an isotropic system, in which the couplings
+$h_{pq}$ depend only on the distance $\abs{p-q}$.  On a $D$
+dimensional lattice, isotropy implies one reflection symmetry for each
+lattice axis, which reduces the degrees of freedom in of
+$\ut{\check{\m h}}_k$ from $\sim N\times N$ to
+$\sim\p{N/2^D}\times\p{N/2^D}$.  We leave this reduction to future
+work.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{The fully collective model}
+
+[TODO: explicitly work out the special case of a fully collective
+model]
 

### 2020-01-23 collective_sun/sun_perturbation.tex
added complete analytical solution for the fully collective model

@@
 enforcing $h_{pq}=h_{qp}$ and $h_{pp}=0$.  We thus find that a
-perturbation $\V_1^\Delta\ket\psi$ indexed by a scalar $\Delta$ is an
+perturbation $\V_1^\Delta\ket\psi$ (indexed by $\Delta$) is an
 eigenvector of the interaction Hamiltonian $H_0$ if its coefficient
@@
 \end{align}
-Interestingly, the energy of the (unnormalized) state
-$\V_1^\Delta\ket\psi$ depends only on the coefficients
-$\v v_X^\Delta$, and is entirely independent of the state $\ket\psi$
-or the choice of a trace-zero single-body operators $X$ used to build
-$\V_1^\Delta$.  Finding operators $\V_1^\Delta$ that generate
-eigenvectors of the interaction Hamiltonian $H_0$ when they act on
-fully symmetric states $\ket\psi$ thus reduces to finding eigenvectors
-of the interaction matrix $\m h-\diag\v h$.
+Interestingly, the energy of the state $\V_1^\Delta\ket\psi$ depends
+only on the coefficients $\v v_X^\Delta$, and is entirely independent
+of the state $\ket\psi$ or the choice of a trace-zero single-body
+operators $X$ used to build $\V_1^\Delta$.  Finding operators
+$\V_1^\Delta$ that generate eigenvectors of the interaction
+Hamiltonian $H_0$ when they act on fully symmetric states $\ket\psi$
+thus reduces to finding eigenvectors of the interaction matrix
+$\m h-\diag\v h$.
 
@@
 enforcing $w_{Opq}=w_{Oqp}$ and $w_{Opp}=0$.  We thus find that the
-vector $\V_2^\Delta\ket\psi$ indexed by a scalar $\Delta$ is an
-eigenvector of the interaction Hamiltonian $H_0$ if its coefficients
+vector $\V_2^\Delta\ket\psi$ (indexed by $\Delta$) is an eigenvector
+of the interaction Hamiltonian $H_0$ if its coefficients
 $\v{\m w}_O^\Delta$ all satisfy the eigenvalue equation
@@
 \end{align}
-which implies that $\V_2^\Delta\ket\psi$ is an (unnormalized state of
-definite energy that depends only on the coefficients $\m w_O^\Delta$.
-Similarly to the case of single-body perturbations, we can decompose
-any vector $\v{\m w}_O$ into its projections $\v{\m w}_O^\Delta$ onto
-the eigenspace of vectors satisfying \eqref{eq:cond_2} with eigenvalue
+which implies that $\V_2^\Delta\ket\psi$ is a state of definite energy
+that depends only on the coefficients $\m w_O^\Delta$.  Similarly to
+the case of single-body perturbations, we can decompose any vector
+$\v{\m w}_O$ into its projections $\v{\m w}_O^\Delta$ onto the
+eigenspace of vectors satisfying \eqref{eq:cond_2} with eigenvalue
 $\Delta$, i.e.~$\v{\m w}_O=\sum_\Delta\v{\m w}_O^\Delta$, which allows
@@
 
-[TODO: explicitly work out the special case of a fully collective
-model]
+Here we solve the single- and two-body eigenvalue problems in the case
+of a fully collective spin model with uniform couplings:
+$h_{pq}=-1+\delta_{pq}$.  In this case, the single-body excitation
+energies $\Delta_k$ are
+\begin{align}
+  \Delta_k
+  = - \sum_{d\ne 0} \sp{\cos\p{d\c k}-1}
+  = N \p{1-\delta_{k,0}}.
+\end{align}
+In order to solve the two-body eigenvalue problem, we expand
+\begin{align}
+  \ut{\check{\m h}}_k
+  = \sum_{d\ne0} 2 \op{d}
+  + \sum_{\substack{c,d\ne0\\c\ne d}} \p{1+e^{-i\p{d-c}\c k}} \op{d}{c},
+\end{align}
+and define the Fourier transform
+\begin{align}
+  F \equiv \sum_{\substack{d\in\ZZ_L^D\\q\in\ZZ_L^D\times2\pi/L}}
+  e^{-id\c q} \op{q}{d},
+\end{align}
+in terms of which
+\begin{align}
+  \f12 F \ut{\check{\m h}}_k F^\dag
+  &= - \f12 N^2 \p{\op{0} + \op{k}} + \sum_{p,q} \op{p}{q} \\
+  &= - \f12 N^2 \p{\op{0} + \op{k}} + \sum_{p,q\in\set{0,k}} \op{p}{q}
+  + \sum_{p,q\notin\set{0,k}} \op{p}{q}
+  + \sum_{\substack{p\in\set{0,k}\\q\notin\set{0,k}}}
+  \p{\op{p}{q} + \op{q}{p}}.
+\end{align}
+Defining the uniform, normalized vector
+\begin{align}
+  \ket{\bar{k}} \equiv \f1{\sqrt{N-2+\delta_{k,0}}}
+  \sum_{p\notin\set{0,k}} \ket{p},
+\end{align}
+we then have
+\begin{multline}
+  \f12 F \ut{\check{\m h}}_k F^\dag
+  = - \f12 N^2 \p{\op{0} + \op{k}}
+  + \sum_{p,q\in\set{0,k}} \op{p}{q} \\
+  + \sqrt{N-2+\delta_{k,0}} \sum_{p\in\set{0,k}}
+  \p{\op{p}{\bar{k}} + \op{\bar{k}}{p}}
+  + \p{N-2+\delta_{k,0}} \op{\bar{k}}.
+\end{multline}
+which is essentially a $2\times 2$ ($3\times 3$) matrix that can be
+diagonalized analytically in the case of $k=0$ ($k\ne0$).  The
+eigenvalues of $F \ut{\check{\m h}}_0 F^\dag$ are
+\begin{align}
+  \varepsilon_0^\pm \equiv \pm N\sqrt{\p{N+3}\p{N-1}} - N\p{N-1},
+\end{align}
+with corresponding (unnormalized) eigenvectors
+\begin{align}
+  \ut{\v{\m w}}_0^\pm
+  = \p{N - 1 - \f12 \Delta_0^\pm} \ket{0} - \sqrt{N-1} \ket{\bar{0}}.
+\end{align}
+Vectors orthogonal to both $\ut{\v{\m w}}_0^\pm$ are eigenvectors of
+$F \ut{\check{\m h}}_0 F^\dag$ with eigenvalue 0.  In the case of
+$k\ne0$, the eigenvalues of $F \ut{\check{\m h}}_{k\ne0} F^\dag$ are
+\begin{align}
+  \varepsilon_k^{(0)} \stackrel{k\ne0}{\equiv} -N^2,
+  &&
+  \varepsilon_k^{(\pm)}
+  \stackrel{k\ne0}{\equiv}
+  \pm \f12 N \sqrt{\p{N+6}\p{N-2}} - \f12 N \p{N-2},
+\end{align}
+with corresponding (unnormalized) eigenvectors
+\begin{align}
+  \ut{\v{\m w}}_k^{(0)}
+  \stackrel{k\ne0}{\equiv} \ket{0} - \ket{k},
+  &&
+  \ut{\v{\m w}}_k^{(\pm)}
+  \stackrel{k\ne0}{\equiv}
+  \p{N - 2 - \f12\Delta_k^{(\pm)}} \p{\ket{0}+\ket{k}}
+  - 2 \sqrt{N-2}\ket{\bar{k}}.
+\end{align}
+In the case of $k\ne0$, vectors orthogonal to all
+$\ut{\v{\m w}}_k^{(s)}$ for $s\in\set{0,+,-}$ are eigenvectors of
+$F \ut{\check{\m h}}_k F^\dag$ with eigenvalue 0.
 

### 2020-01-23 collective_sun/sun_perturbation.tex
fixed error in analytics for fully collective model

@@
   \m m_{dk} \equiv \sum_p e^{ip\c k} \op{p+d}{p},
+  \label{eq:cycle}
 \end{align}
@@
   \ut{\v{\m w}}_k \equiv \sum_{d\ne0} \tilde w_{dk} \ket{d},
+  \label{eq:cycle_vecs}
 \end{align}
@@
   \ut{\check{\m h}}_k
-  = \sum_{d\ne0} 2 \op{d}
-  + \sum_{\substack{c,d\ne0\\c\ne d}} \p{1+e^{-i\p{d-c}\c k}} \op{d}{c},
-\end{align}
-and define the Fourier transform
-\begin{align}
-  F \equiv \sum_{\substack{d\in\ZZ_L^D\\q\in\ZZ_L^D\times2\pi/L}}
-  e^{-id\c q} \op{q}{d},
+  = - \sum_{d\ne0} 2 \op{d}
+  - \sum_{\substack{c,d\ne0\\c\ne d}} \p{1+e^{-i\p{d-c}\c k}} \op{d}{c}
+  = - \sum_{c,d\ne0} \p{1+e^{-i\p{d-c}\c k}} \op{d}{c},
+\end{align}
+and define the vectors
+\begin{align}
+  \ket{\ut{p}}
+  \equiv \f1{\sqrt{N-1}} \sum_{d\ne 0} e^{i p\c d}\ket{d},
 \end{align}
@@
 \begin{align}
-  \f12 F \ut{\check{\m h}}_k F^\dag
-  &= - \f12 N^2 \p{\op{0} + \op{k}} + \sum_{p,q} \op{p}{q} \\
-  &= - \f12 N^2 \p{\op{0} + \op{k}} + \sum_{p,q\in\set{0,k}} \op{p}{q}
-  + \sum_{p,q\notin\set{0,k}} \op{p}{q}
-  + \sum_{\substack{p\in\set{0,k}\\q\notin\set{0,k}}}
-  \p{\op{p}{q} + \op{q}{p}}.
-\end{align}
-Defining the uniform, normalized vector
-\begin{align}
-  \ket{\bar{k}} \equiv \f1{\sqrt{N-2+\delta_{k,0}}}
-  \sum_{p\notin\set{0,k}} \ket{p},
-\end{align}
-we then have
-\begin{multline}
-  \f12 F \ut{\check{\m h}}_k F^\dag
-  = - \f12 N^2 \p{\op{0} + \op{k}}
-  + \sum_{p,q\in\set{0,k}} \op{p}{q} \\
-  + \sqrt{N-2+\delta_{k,0}} \sum_{p\in\set{0,k}}
-  \p{\op{p}{\bar{k}} + \op{\bar{k}}{p}}
-  + \p{N-2+\delta_{k,0}} \op{\bar{k}}.
-\end{multline}
-which is essentially a $2\times 2$ ($3\times 3$) matrix that can be
-diagonalized analytically in the case of $k=0$ ($k\ne0$).  The
-eigenvalues of $F \ut{\check{\m h}}_0 F^\dag$ are
-\begin{align}
-  \varepsilon_0^\pm \equiv \pm N\sqrt{\p{N+3}\p{N-1}} - N\p{N-1},
-\end{align}
-with corresponding (unnormalized) eigenvectors
-\begin{align}
-  \ut{\v{\m w}}_0^\pm
-  = \p{N - 1 - \f12 \Delta_0^\pm} \ket{0} - \sqrt{N-1} \ket{\bar{0}}.
-\end{align}
-Vectors orthogonal to both $\ut{\v{\m w}}_0^\pm$ are eigenvectors of
-$F \ut{\check{\m h}}_0 F^\dag$ with eigenvalue 0.  In the case of
-$k\ne0$, the eigenvalues of $F \ut{\check{\m h}}_{k\ne0} F^\dag$ are
-\begin{align}
-  \varepsilon_k^{(0)} \stackrel{k\ne0}{\equiv} -N^2,
-  &&
-  \varepsilon_k^{(\pm)}
-  \stackrel{k\ne0}{\equiv}
-  \pm \f12 N \sqrt{\p{N+6}\p{N-2}} - \f12 N \p{N-2},
-\end{align}
-with corresponding (unnormalized) eigenvectors
-\begin{align}
-  \ut{\v{\m w}}_k^{(0)}
-  \stackrel{k\ne0}{\equiv} \ket{0} - \ket{k},
-  &&
-  \ut{\v{\m w}}_k^{(\pm)}
-  \stackrel{k\ne0}{\equiv}
-  \p{N - 2 - \f12\Delta_k^{(\pm)}} \p{\ket{0}+\ket{k}}
-  - 2 \sqrt{N-2}\ket{\bar{k}}.
-\end{align}
-In the case of $k\ne0$, vectors orthogonal to all
-$\ut{\v{\m w}}_k^{(s)}$ for $s\in\set{0,+,-}$ are eigenvectors of
-$F \ut{\check{\m h}}_k F^\dag$ with eigenvalue 0.
+  \ut{\check{\m h}}_k
+  = -\p{N-1} \p{\op{\ut{0}} + \op{\ut{k}}}.
+\end{align}
+The vectors $\ket{\ut{p}}$ are normalized, but not orthogonal:
+$\braket{\ut{p}|\ut{q}}=\delta_{pq}-\p{1-\delta_{pq}}/\p{N-1}$.  We
+therefore have
+\begin{align}
+  \ut{\check{\m h}}_k
+  = - \sum_{s\in\set{\pm}} \varepsilon_s \op{\ut{k}^s}
+\end{align}
+where
+\begin{align}
+  \varepsilon_\pm \equiv N - 1 \mp 1,
+  &&
+  \ket{\ut{k}^\pm}
+  \equiv \p{1\mp\f1{N-1}}^{-1/2}
+  \f{\ket{\ut{0}}\pm\ket{\ut{k}}}{\sqrt{2}}.
+\end{align}
+The vectors $\ket{\ut{k}^\pm}$ are orthonormal when $k\ne0$.  When
+$k=0$, meanwhile, we have
+\begin{align}
+  \ut{\check{\m h}}_0 = -2\p{N-1} \op{\ut{0}}.
+\end{align}
+Remembering from \eqref{eq:cycle_vecs} that vectors $\ket{\ut{k}^\pm}$
+are associated with the cycle matrices $\m m_{dk}$ defined in
+\eqref{eq:cycle} via
+\begin{align}
+  \ket{\ut{k}^\pm}
+  \propto \ket{\ut{0}} \pm \ket{\ut{k}}
+  \propto \sum_{d\ne0} \p{1 \pm e^{ik\c d}} \ket{d}
+  \to \sum_{d\ne0} \p{1 \pm e^{ik\c d}} \m m_{dk},
+\end{align}
+one can work out that the vectors $\ket{\ut{k}^\pm}$ correspond to
+matrices that are strictly symmetric ($+$) or anti-symmetric ($-$)
+under transposition.  The two-body perturbations considered in this
+work, however, are defined using coefficient vectors
+$\set{\v{\m w}_O}$ that correspond to symmetric matrices
+$\set{\m w_O}$, with $w_{Opq}=w_{Oqp}$.  We can therefore neglect the
+eigenvectors $\ket{\ut{k}^-}$ and corresponding eigenvalues
+$\varepsilon_-$ entirely for the purposes of this work.
+
+Altogether, the two-body excitation energies corresponding to
+non-trivial eigenstates of $\ut{\check{\m h}}_k$ are
+\begin{align}
+  \Delta_0 \equiv -2\p{N-1} - 2 h_0 = 0,
+  &&
+  \Delta_+ \equiv -\varepsilon_+ - 2 h_0 = N,
+\end{align}
+while the energy of all orthogonal two-body excitations is
+\begin{align}
+  \Delta_\perp \equiv - 2 h_0 = 2N - 2.
+\end{align}
 

### 2020-01-24 collective_sun/sun_perturbation.tex
better formalize the use of integers modulo L

@@
 
-We consider an array of $N$ multilevel spins with non-uniform
-SU($n$)-symmetric interactions that can be written in the form
+We consider an array of $N$ multilevel spins with SU($n$)-symmetric
+interactions that can be written in the form
 \begin{align}
@@
 multilevel generalization of the SU(2)-symmetric spin interaction
-$\v S_p\c\v S_q$ for $\v S=\p{S_\x,S_\y,S_\z}$.
+$\v S_p\c\v S_q$ for $\v S=\p{s_\x,s_\y,s_\z}$, with $s_\alpha$ an
+individual spin operator proportional to the Pauli matrix
+$\sigma_\alpha$.
 
@@
   \Delta
-  = \sum_{d\in\tilde\ZZ_L^D} h_{0,d} \sp{\cos\p{d \c k_{\t{SE}}}-1}
-  = h \sum_{\substack{d\in\tilde\ZZ_L^D\\d\ne0}}
+  = \sum_{d\in\ZZ_L^D} h_{0,d} \sp{\cos\p{d \c k_{\t{SE}}}-1}
+  = h \sum_{\substack{d\in\ZZ_L^D\\\abs{d}\ge1}}
   \f{1-\cos\p{d \c k_{\t{SE}}}}{\abs{d}^\alpha},
 \end{align}
-where the domain of $d$ is given by
-$\tilde\ZZ_L\equiv\set{-\floor{L/2},-\floor{L/2}+1,\cdots,\ceil{L/2}}$,
-and $k_{\t{SE}}$ is a wavenumber for a singly-excited spin-wave state.
-In order to yield the smallest excitation energy $\Delta$, the
-wavenumber $k_{\t{SE}}$ should maximize the contribution of the cosine
-term above, which is achieved by a wavenumber that minimizes the
-oscillations of this term when integrated over the entire lattice.  A
-suitable candidate for a minimal wavenumber is
-$k_{\t{SE}}=\p{2\pi/L,0,0,\cdots}$, which corresponds to an excitation
-energy
-\begin{align}
-  \Delta = h \sum_{\substack{d\in\tilde\ZZ_L^D\\d\ne0}}
+where the domain of $d$ is defined in terms of integers modulo $L$,
+i.e.~$\ZZ_L\simeq\set{0,1,\cdots,L-1}$ with the relation $\simeq$
+denoting an association that ignores the cyclic structure of $\ZZ_L$;
+$k_{\t{SE}}\in\ZZ_L\times2\pi/L$ is a wavenumber for a singly-excited
+spin-wave state; and the norm $\abs{d}$ for a vector $d$ on a periodic
+lattice is implicitly understood to mean the smallest Euclidean
+distance of $d$ from a fixed origin.  In order to yield the smallest
+excitation energy $\Delta$, the wavenumber $k_{\t{SE}}$ should
+maximize the contribution of the cosine term above, which is achieved
+by a wavenumber that minimizes the oscillations of this term when
+integrated over the entire lattice.  A suitable candidate for a
+minimal wavenumber is $k_{\t{SE}}=\p{2\pi/L,0,0,\cdots}$, which
+corresponds to an excitation energy
+\begin{align}
+  \Delta = h \sum_{\substack{d\in\ZZ_L^D\\\abs{d}\ge1}}
   \f{1-\cos\p{d_12\pi/L}}{\abs{d}^\alpha}.
 \end{align}
-Defining $\epsilon\equiv2/L$ and a rescaled domain
-$\SS_\epsilon\equiv\tilde\ZZ_L/\epsilon\subset\sp{-1,1}$, we
-substitute $x\equiv\epsilon d\in\SS_\epsilon^D$ to get
+Defining $\epsilon\equiv2/L$ and a rescaled domain symmetrized about
+$0$, $\SS_\epsilon\simeq\ZZ_L/\epsilon$ with
+$\SS_\epsilon\subset\sp{-1,1}$, we substitute $x\simeq\epsilon d$ to
+get
 \begin{align}
   \Delta
-  = h \sum_{\substack{x\in\SS_\epsilon^D\\x\ne0}}
+  = h \sum_{\substack{x\in\SS_\epsilon^D\\\abs{x}\ge\epsilon}}
   \f{1-\cos\p{\pi x_1}}{\abs{x/\epsilon}^\alpha}
-  = h \epsilon^{D-\alpha} \sum_{\substack{x\in\SS_L^D\\x\ne0}}
+  = h \epsilon^{\alpha-D} \sum_{\substack{x\in\SS_L^D\\\abs{x}\ge\epsilon}}
   \epsilon^D \f{1-\cos\p{\pi x_1}}{\abs{x}^\alpha}.
@@
 sites are indexed by vectors
-$p\in\tilde\ZZ_L^D \equiv
-\set{-\floor{L/2},-\floor{L/2}+1,\cdots,\ceil{L/2}}$, and wavenumbers
-take on values $k\in\tilde\ZZ_L^D\times2\pi/L$.  The corresponding
-eigenvalues of $\m h$ can be determined by expanding
+$p\in\ZZ_L^D\simeq\set{0,1,\cdots,L-1}^D$, and wavenumbers take on
+values $k\in\ZZ_L^D\times2\pi/L$.  The corresponding eigenvalues of
+$\m h$ can be determined by expanding
 \begin{align}
@@
 translational symmetry of $\m h$.  To this end, for all lattice
-displacements $d\in\tilde\ZZ_L^D$ and wavenumbers
-$k\in\tilde\ZZ_L^D\times2\pi/L$ we define the {\it cycle matrices}
+displacements $d\in\ZZ_L^D$ and wavenumbers $k\in\ZZ_L^D\times2\pi/L$
+we define the {\it cycle matrices}
 \begin{align}

### 2020-01-26 collective_sun/sun_perturbation.tex
minor change to language
simplified some math

@@
   = E_0 \V_2 \ket\psi
-  + \f12 \sum_{p\ne q} \sp{2\bar h_{pq} w_{Opq}
+  + \f12 \sum_O \sum_{p\ne q} \sp{2\bar h_{pq} w_{Opq}
     + \sum_k \p{h_{pk} w_{Okq} + w_{Opk} h_{kq}}}
@@
   H_0 \V_1 \ket\psi
-  = \sum_{X,k} \sum_{p<q} h_{pq} v_{Xk} \Pi_{pq} X_k \ket\psi,
+  = \sum_k \sum_{p<q} h_{pq} v_k \Pi_{pq} X_k \ket\psi,
   \label{eq:H_V1_psi_start}
 \end{align}
-which has terms with $k\in\set{p,q}$, and terms with
-$k\notin\set{p,q}$.  In the case of $k\notin\set{p,q}$, the
-permutation operator $\Pi_{pq}$ commutes with $X_k$ and annihilates on
-$\ket\psi$, leaving terms at each fixed $k$ of the form
+where without loss of generality we consider a single-body
+perturbation $\V_1$ built out of only one single-body operator $X$.
+The sum in \eqref{eq:H_V1_psi_start} has terms with $k\in\set{p,q}$,
+and terms with $k\notin\set{p,q}$.  In the case of $k\notin\set{p,q}$,
+the permutation operator $\Pi_{pq}$ commutes with $X_k$ and
+annihilates on $\ket\psi$, leaving terms at each fixed $k$ of the form
 \begin{align}
@@
 \begin{align}
-  \sum_X \sum_{\substack{p<q\\k\notin\set{p,q}}}
-  h_{pq} v_{Xk} \Pi_{pq} X_k \ket\psi
-  = \sum_{X,k} \p{E_0 - h_k} v_{Xk} X_k \ket\psi
-  = E_0 \V_1 \ket\psi - \sum_{X,k} h_k v_{Xk} X_k \ket\psi
-\end{align}
-while the terms in \eqref{eq:H_V1_psi_start} with $k\in\set{p,q}$ for
-each fixed $X$ take the form
+  \sum_{\substack{p<q\\k\notin\set{p,q}}}
+  h_{pq} v_k \Pi_{pq} X_k \ket\psi
+  = \sum_k \p{E_0 - h_k} v_k X_k \ket\psi
+  = E_0 \V_1 \ket\psi - \sum_k h_k v_k X_k \ket\psi
+\end{align}
+while the terms in \eqref{eq:H_V1_psi_start} with $k\in\set{p,q}$ take
+the form
 \begin{align}
   \sum_{\substack{p<q\\k\in\set{p,q}}}
-  h_{pq} v_{Xk} \Pi_{pq} X_k \ket\psi
-  = \sum_{k<q} h_{kq} v_{Xk} X_q \ket\psi
-  + \sum_{p<k} h_{pk} v_{Xk} X_p \ket\psi
-  = \sum_{k,q} h_{kq} v_{Xk} X_q \ket\psi
+  h_{pq} v_k \Pi_{pq} X_k \ket\psi
+  = \sum_{k<q} h_{kq} v_k X_q \ket\psi
+  + \sum_{p<k} h_{pk} v_k X_p \ket\psi
+  = \sum_{k,q} h_{kq} v_k X_q \ket\psi
 \end{align}
@@
   = E_0 \V_1 \ket\psi
-  + \sum_X \p{\sum_{p,q} h_{pq} v_{Xp} X_q - \sum_p h_p v_{Xp} X_p}
+  + \p{\sum_{p,q} h_{pq} v_p X_q - \sum_p h_p v_p X_p}
   \ket\psi.
@@
   H_0 \V_2 \ket\psi
-  = \sum_O \sum_{\substack{k<\ell\\p<q}} w_{Ok\ell} h_{pq}
+  = \sum_{\substack{k<\ell\\p<q}} w_{k\ell} h_{pq}
   \Pi_{pq} O_{k\ell} \ket\psi,
@@
 \end{align}
-which has terms with $\set{p,q}=\set{k,\ell}$, terms with
-$p,q\notin\set{k,\ell}$, and terms with only one of $p$ or
+where without loss of generality we consider a two-body perturbation
+$\V_2$ built out of only one two-body operator $O$.  The sum in
+\eqref{eq:H_V2_psi_start} has terms with $\set{p,q}=\set{k,\ell}$,
+terms with $p,q\notin\set{k,\ell}$, and terms with only one of $p$ or
 $q\in\set{k,\ell}$.  The permutation operator $\Pi_{pq}$ acts
-trivially on $O_{k_\ell}\ket\psi$ when $\set{p,q}=\set{k,\ell}$ or
+trivially on $O_{k\ell}\ket\psi$ when $\set{p,q}=\set{k,\ell}$ or
 $p,q\notin\set{k,\ell}$, leaving terms at each fixed $k,\ell$ of the
@@
   E_0 \V_2 \ket\psi
-  + \sum_O \sum_{k<\ell} 2 \bar h_{k\ell} w_{Ok\ell} O_{k\ell} \ket\psi.
+  + \sum_{k<\ell} 2 \bar h_{k\ell} w_{k\ell} O_{k\ell} \ket\psi.
 \end{align}
@@
 \begin{align}
-  \f12 \sum_{k\ne\ell} \sum_q \p{h_{kq} w_{Oq\ell} + w_{Okq} h_{q\ell}}
+  \f12 \sum_{k\ne\ell} \sum_q \p{h_{kq} w_{q\ell} + w_{kq} h_{q\ell}}
   O_{k\ell} \ket\psi.
@@
   &= E_0 \V_2 \ket\psi
-  + \sum_{k<\ell} 2\bar h_{k\ell} w_{Ok\ell} O_{k\ell} \ket\psi
-  + \f12 \sum_{k\ne\ell} \sum_q \p{h_{kq} w_{Oq\ell} + w_{Okq} h_{q\ell}}
+  + \sum_{k<\ell} 2\bar h_{k\ell} w_{k\ell} O_{k\ell} \ket\psi
+  + \f12 \sum_{k\ne\ell} \sum_q \p{h_{kq} w_{q\ell} + w_{kq} h_{q\ell}}
   O_{k\ell} \ket\psi \\
   &= E_0 \V_2 \ket\psi
-  + \f12 \sum_{k\ne\ell} \sp{2\bar h_{k\ell} w_{Ok\ell}
-    + \sum_q \p{h_{kq} w_{Oq\ell} + w_{Okq} h_{q\ell}}}
+  + \f12 \sum_{k\ne\ell} \sp{2\bar h_{k\ell} w_{k\ell}
+    + \sum_q \p{h_{kq} w_{q\ell} + w_{Okq} h_{q\ell}}}
   O_{k\ell} \ket\psi.
@@
 
-Altogether, the two-body excitation energies corresponding to
-non-trivial eigenstates of $\ut{\check{\m h}}_k$ are
+Altogether, the excitation energies corresponding to non-trivial
+eigenstates of $\ut{\check{\m h}}_k$ are
 \begin{align}

### 2020-01-26 collective_sun/sun_perturbation.tex
added a calculation of the general M-body energy eigenvalue problem

@@
 \newcommand{\B}{\mathcal{B}}
+\newcommand{\C}{\mathcal{C}}
 \newcommand{\D}{\mathcal{D}}
@@
 $H_2^{(2)}$, as before we pick an arbitrary fully symmetric state
-$\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_V2_psi})
+$\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_VM_psi})
 \begin{align}
@@
   = E_0 \V_2 \ket\psi
-  + \f12 \sum_O \sum_{p\ne q} \sp{2\bar h_{pq} w_{Opq}
-    + \sum_k \p{h_{pk} w_{Okq} + w_{Opk} h_{kq}}}
+  + \sum_O \sum_{p<q} \sp{\bar h_{pq} w_{Opq}
+    + \sum_k \p{h_{kp} w_{Okq} w_{Okq} + h_{kq} w_{Opk}}}
   O_{pq} \ket\psi,
@@
 \begin{align}
-  \bar h_{pq} \equiv h_{pq} - \f12\p{h_p + h_q},
+  \bar h_{pq} \equiv 2 h_{pq} - h_p - h_q,
   &&
@@
 \begin{align}
-  \check{\m h} \equiv 2\sum_{p<q} \bar h_{pq} \op{\set{p,q}}
+  \check{\m h} \equiv \sum_{p<q} \bar h_{pq} \op{\set{p,q}}
   + \sum_{p<q} \ket{\set{p,q}} \sum_{k\notin\set{p,q}}
-  \p{h_{pk} \bra{\set{k,q}} + h_{kq} \bra{\set{p,k}}}.
+  \p{h_{kp} \bra{\set{k,q}} + h_{kq} \bra{\set{p,k}}}.
   \label{eq:h_super_mat}
@@
 \end{align}
-where without loss of generality we consider a single-body
-perturbation $\V_1$ built out of only one single-body operator $X$.
-The sum in \eqref{eq:H_V1_psi_start} has terms with $k\in\set{p,q}$,
-and terms with $k\notin\set{p,q}$.  In the case of $k\notin\set{p,q}$,
-the permutation operator $\Pi_{pq}$ commutes with $X_k$ and
-annihilates on $\ket\psi$, leaving terms at each fixed $k$ of the form
+where $\ket\psi\in\M_\FS$ is an arbitrary fully symmetric state, and
+without loss of generality we consider a single-body perturbation
+$\V_1$ built out of only one single-body operator $X$.  The sum in
+\eqref{eq:H_V1_psi_start} has terms with $k\in\set{p,q}$, and terms
+with $k\notin\set{p,q}$.  In the case of $k\notin\set{p,q}$, the
+permutation operator $\Pi_{pq}$ commutes with $X_k$ and annihilates on
+$\ket\psi$, leaving terms at each fixed $k$ of the form
 \begin{align}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Two-body perturbation}
-\label{sec:H_V2_psi}
-
-We wish to simplify
-\begin{align}
-  H_0 \V_2 \ket\psi
-  = \sum_{\substack{k<\ell\\p<q}} w_{k\ell} h_{pq}
-  \Pi_{pq} O_{k\ell} \ket\psi,
-  \label{eq:H_V2_psi_start}
-\end{align}
-where without loss of generality we consider a two-body perturbation
-$\V_2$ built out of only one two-body operator $O$.  The sum in
-\eqref{eq:H_V2_psi_start} has terms with $\set{p,q}=\set{k,\ell}$,
-terms with $p,q\notin\set{k,\ell}$, and terms with only one of $p$ or
-$q\in\set{k,\ell}$.  The permutation operator $\Pi_{pq}$ acts
-trivially on $O_{k\ell}\ket\psi$ when $\set{p,q}=\set{k,\ell}$ or
-$p,q\notin\set{k,\ell}$, leaving terms at each fixed $k,\ell$ of the
+\subsection{Multi-body perturbation}
+\label{sec:H_VM_psi}
+
+Given an $M$-body perturbation
+\begin{align}
+  \V_M \equiv \sum_{k\in\C_N\p{M}} w_k O_k,
+\end{align}
+where $O$ is an $M$-body operator; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is
+a choice of $M$ distinct spins that $O$ acts on; and $\C_N\p{n}$ is a
+choice of $n$ elements from a set of $N$; we wish to simplify
+\begin{align}
+  H_0 \V_M \ket\psi
+  = \sum_{\substack{\p{p,q}\in\C_N\p{2}\\k\in\C_N\p{M}}} h_{pq} w_k
+  \Pi_{pq} O_k \ket\psi,
+  \label{eq:H_VM_psi_start}
+\end{align}
+where $\ket\psi\in\M_\FS$ is an arbitrary fully symmetric state.  The
+sum in \eqref{eq:H_VM_psi_start} has terms with $p,q\in k$, terms with
+$p,q\notin k$, and terms with only one of $p$ or $q\in k$.  The
+permutation operator $\Pi_{pq}$ acts trivially on $O_k\ket\psi$ when
+$p,q\in k$ or $p,q\notin k$, leaving terms at each fixed $k$ of the
 form
 \begin{align}
-  h_{k\ell} + \sum_{\substack{p<q\\p,q\notin\set{k,\ell}}} h_{pq}
-  = \sum_{p<q} h_{pq}
-  - \sum_{\substack{p\in\set{k,\ell}\\q\notin\set{k,\ell}}} h_{pq}
-  = E_0 - h_k - h_\ell + 2 h_{k\ell}
-  = E_0 + 2 \bar h_{k\ell},
-\end{align}
-where
-\begin{align}
-  \bar h_{k\ell} \equiv h_{k\ell} - \f12\p{h_k + h_\ell}.
-\end{align}
-The corresponding terms in \eqref{eq:H_V2_psi_start} with
-$\set{p,q}=\set{k,\ell}$ or $p,q\notin\set{k,\ell}$ are then
-\begin{align}
-  E_0 \V_2 \ket\psi
-  + \sum_{k<\ell} 2 \bar h_{k\ell} w_{k\ell} O_{k\ell} \ket\psi.
-\end{align}
-Using similar techniques, we can simplify the terms in
-\eqref{eq:H_V2_psi_start} with only one of $p,q\in\set{k,\ell}$ to
-\begin{align}
-  \f12 \sum_{k\ne\ell} \sum_q \p{h_{kq} w_{q\ell} + w_{kq} h_{q\ell}}
-  O_{k\ell} \ket\psi.
-\end{align}
-In total,
-\begin{align}
-  H_0 \V_2 \ket\psi
-  &= E_0 \V_2 \ket\psi
-  + \sum_{k<\ell} 2\bar h_{k\ell} w_{k\ell} O_{k\ell} \ket\psi
-  + \f12 \sum_{k\ne\ell} \sum_q \p{h_{kq} w_{q\ell} + w_{kq} h_{q\ell}}
-  O_{k\ell} \ket\psi \\
-  &= E_0 \V_2 \ket\psi
-  + \f12 \sum_{k\ne\ell} \sp{2\bar h_{k\ell} w_{k\ell}
-    + \sum_q \p{h_{kq} w_{q\ell} + w_{Okq} h_{q\ell}}}
-  O_{k\ell} \ket\psi.
+  \sum_{\substack{\p{p,q}\in\C_N\p{2}\\p,q\in k~\t{or}~p,q\notin k}} h_{pq}
+  = \sum_{\p{p,q}\in\C_N\p{2}} h_{pq}
+  - \sum_{p\in k} \sum_{\substack{q\in\ZZ_N\\q\notin k}} h_{pq}
+  = E_0 - \sum_{p\in k}
+  \p{\sum_{q\in\ZZ_N} h_{pq} - \sum_{q\in k} h_{pq}},
+\end{align}
+where we first split a sum over $\p{p,q}\in\C_N\p{2}$ with the
+restriction that $p,q\in k~\t{or}~p,q\notin k$ into an unrestricted
+sum minus a remainder, and then similarly split the sum over all spins
+$q$ with the restriction that $q\notin k$ into a sum over all $q$
+minus the remainder $q\in k$.  Defining $h_p\equiv\sum_q h_{pq}$, we thus find that
+\begin{align}
+  \sum_{\substack{\p{p,q}\in\C_N\p{2}\\p,q\in k~\t{or}~p,q\notin k}} h_{pq}
+  = E_0 + 2\sum_{\p{p,q}\subset k} h_{pq} - \sum_{p\in k} h_p,
+\end{align}
+where the sum over $\p{p,q}\subset k$ denotes a sum over distinct
+choices of two elements in $k$.  The terms in
+\eqref{eq:H_VM_psi_start} with only one of $p\in k$ or $q\in k$,
+meanwhile, are
+\begin{align}
+  \sum_{k\in\C_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}} \sum_{q\in k}
+  h_{pq} w_k \Pi_{pq} O_k \ket\psi
+  = \sum_{k\in\C_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}}
+  \sum_{j=1}^M h_{p k_j} w_k \Pi_{p k_j} O_k \ket\psi.
+\end{align}
+Defining
+$k_{jp}\equiv\p{k_1,k_2,\cdots,k_{j-1},p,k_{j+1},\cdots,k_M}$, i.e.~an
+element of $\C_N\p{M}$ in which the $j$-th number $k_j$ has been
+replaced by $p$, we can simplify
+\begin{align}
+  \Pi_{p k_j} O_k \ket\psi = O_{k_{jp}}\ket\psi,
+\end{align}
+and switch the labels for $p$ and $k_j$ to find that
+\begin{align}
+  \sum_{k\in\C_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}} \sum_{q\in k}
+  h_{pq} w_k  \Pi_{pq} O_k \ket\psi
+  = \sum_{k\in\C_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}}
+  \sum_{j=1}^M h_{pk_j} w_{k_{jp}} O_k \ket\psi.
+\end{align}
+Defining
+\begin{align}
+  \p{h\circ w}_k
+  \equiv \sum_{j=1}^M \sum_{p\in k} h_{pk_j} w_{k_{jp}}
+  = h_{pk_1} w_{pk_2k_3\cdots k_M}
+  + h_{pk_2} w_{k_1pk_3\cdots k_M}
+  + h_{pk_M} \cdots w_{k_1k_2\cdots k_{M-1} p}
+\end{align}
+and
+\begin{align}
+  \bar h_k \equiv 2 \sum_{\p{p,q}\subset k} h_{pq} - \sum_{p\in k} h_p,
+\end{align}
+we thus find that
+\begin{align}
+  H_0 \V_M \ket\psi
+  = E_0 \V_M \ket\psi + \sum_{k\in\C_N\p{M}}
+  \sp{\bar h_k w_k + \p{h\circ w}_k} O_k \ket\psi.
 \end{align}
@@
 \begin{align}
-  2 \m h\circ \m w + \P_{\t{off-diag}}\p{\m h\c\m w + \m w\c\m h}
+  2 \m h*\m w + \P_{\t{off-diag}}\p{\m h\c\m w + \m w\c\m h}
   = \p{\Delta + 2h_0} \m w,
@@
 \end{align}
-where $\m h$ and $\m w$ are matrices of all $h_{pq}$ and $w_{pq}$, and
-$\m h\circ\m w$ denotes an element-wise (Kronecker) product of $\m h$
-and $\m w$, and
+where $\m h$ and $\m w$ are matrices of all $h_{pq}$ and $w_{pq}$;
+$\m h*\m w$ denotes an element-wise (Kronecker) product of $\m h$ and
+$\m w$, i.e.~with matrix elements $h_{pq}w_{pq}$; and
 $\P_{\t{off-diag}}\p{\m m}\equiv \m m-\sum_p\op{p}\m m\op{p}$ denotes

### 2020-01-26 collective_sun/sun_perturbation.tex
some improvements to notation in the two-body eigenvalue problem in the appendices

@@
 \begin{align}
-  \p{2h_{pq}-h_p-h_q} w_{pq}
+  \bar h_{pq} w_{pq}
   + \sum_k \p{h_{pk} w_{kq} + w_{pk} h_{kq}}
   = \Delta w_{pq},
+  &&
+  \bar h_{pq} \equiv 2h_{pq} - h_p - h_q,
+  &&
+  h_p \equiv \sum_q h_{pq},
   \label{eq:cond_2_ref_start}
 \end{align}
-where $h_p\equiv\sum_q h_{pq}$.  If the interaction Hamiltonian $H_0$
-is translationally invariant, then the couplings $h_{pq}$ depend on
-$p$ and $q$ only through the separation $\pm\p{p-q}$.  In this case,
-all $h_p=h_0$ are equal, and \eqref{eq:cond_2_ref_start} can be
-written in the form
-\begin{align}
-  2 \m h*\m w + \P_{\t{off-diag}}\p{\m h\c\m w + \m w\c\m h}
-  = \p{\Delta + 2h_0} \m w,
+which can be written in matrix form as
+\begin{align}
+  \bar{\m h}*\m w + \P_{\t{off-diag}}\p{\m h\c\m w + \m w\c\m h}
+  = \Delta \m w,
   \label{eq:cond_2_ref}
 \end{align}
-where $\m h$ and $\m w$ are matrices of all $h_{pq}$ and $w_{pq}$;
-$\m h*\m w$ denotes an element-wise (Kronecker) product of $\m h$ and
-$\m w$, i.e.~with matrix elements $h_{pq}w_{pq}$; and
+where $\m h$, $\bar{\m h}$, and $\m w$ are matrices of all $h_{pq}$,
+$\bar h_{pq}$ and $w_{pq}$; $\bar{\m h}*\m w$ denotes an element-wise
+(Kronecker) product of $\bar{\m h}$ and $\m w$, i.e.~with matrix
+elements $\bar h_{pq}w_{pq}$; and
 $\P_{\t{off-diag}}\p{\m m}\equiv \m m-\sum_p\op{p}\m m\op{p}$ denotes
 the projection of $\m m$ onto the space of strictly off-diagonal
-matrices.  In order to find solutions to \eqref{eq:cond_2_ref}, we we
-expand the matrices $\m h,\m w$ in a basis that the respects
-translational symmetry of $\m h$.  To this end, for all lattice
-displacements $d\in\ZZ_L^D$ and wavenumbers $k\in\ZZ_L^D\times2\pi/L$
-we define the {\it cycle matrices}
+matrices.  In order to find solutions to \eqref{eq:cond_2_ref} when
+the interaction Hamiltonian $H_0$ is translationally invariant, in
+which case $h_{pq}$ and $\bar h_{pq}$ depend only on the separation
+$\pm\p{p-q}$, we expand all matrices in a basis that the respects
+translational symmetry.  To this end, for all lattice displacements
+$d\in\ZZ_L^D$ and wavenumbers $k\in\ZZ_L^D\times2\pi/L$ we define the
+{\it cycle matrices}
 \begin{align}
@@
 where the element-wise sums in $p+d$ are implicitly taken modulo $L$
-(i.e.~the length of the periodic lattice along each axis).  In terms of the cycle matrices, we define
+(i.e.~the length of the periodic lattice along each axis).  In terms
+of these cycle matrices, we define
 \begin{align}
@@
 \end{align}
-expand
-\begin{align}
-  \m h = \sum_{p,q} h_{pq} \op{p}{q}
-  = \sum_{q,d} h_{q+d,q} \op{q+d}{q}
+and expand
+\begin{align}
+  \m w = \sum_{d,k} \tilde w_{dk} \m m_{dk},
+  &&
+  \m h = \sum_{q,d} h_{q+d,q} \op{q+d}{q}
   = \sum_d h_{d,0} \m m_{d,0},
   &&
-  \m w = \sum_{d,k} \tilde w_{dk} \m m_{dk},
-\end{align}
-and write the condition in \eqref{eq:cond_2_ref} as a new set of
-conditions
-\begin{align}
-  \sum_{d\ne0} 2 h_{d,0} \tilde w_{dk} \m m_{dk}
+  \bar{\m h} = \sum_d \bar h_{d,0} \m m_{d,0}.
+\end{align}
+We can then write the condition in \eqref{eq:cond_2_ref} as a new set
+of conditions
+\begin{align}
+  \sum_{d\ne0} \bar h_{d,0} \tilde w_{dk} \m m_{dk}
   + \sum_{\substack{c,d\\c+d\ne0}}
   h_{c,0} \tilde w_{dk} \p{1 + e^{-ic\c k}} \m m_{c+d,k}
-  = \p{\Delta + 2 h_0} \sum_{d\ne0} \tilde w_{dk} \m m_{dk},
+  = \Delta \sum_{d\ne0} \tilde w_{dk} \m m_{dk},
 \end{align}
@@
 \begin{align}
-  \sum_{d\ne0} 2 h_{d,0} \tilde w_{dk} \m m_{dk}
+  \sum_{d\ne0} \bar h_{d,0} \tilde w_{dk} \m m_{dk}
   + \sum_{c,d\ne0} h_{d-c,0} \tilde w_{ck}
   \p{1 + e^{-i\p{d-c}\c k}} \m m_{dk}
-  = \p{\Delta + 2 h_0} \sum_{d\ne0} \tilde w_{dk} \m m_{dk},
+  = \Delta \sum_{d\ne0} \tilde w_{dk} \m m_{dk},
   \label{eq:cond_2_cycle}
@@
   \ut{\check{\m h}}_k
-  \equiv \sum_{d\ne0} 2 h_{d,0} \op{d}
+  \equiv \sum_{d\ne0} \bar h_{d,0} \op{d}
   + \sum_{c,d\ne0} h_{d-c,0} \p{1+e^{-i\p{d-c}\c k}} \op{d}{c},
@@
 \begin{align}
-  \ut{\check{\m h}}_k \c \ut{\v{\m w}}_k
-  = \p{\Delta + 2 h_0} \ut{\v{\m w}}_k.
+  \ut{\check{\m h}}_k \c \ut{\v{\m w}}_k = \Delta \ut{\v{\m w}}_k.
   \label{eq:cond_2_square}
@@
   \ut{\check{\m h}}_k
-  = - \sum_{d\ne0} 2 \op{d}
-  - \sum_{\substack{c,d\ne0\\c\ne d}} \p{1+e^{-i\p{d-c}\c k}} \op{d}{c}
-  = - \sum_{c,d\ne0} \p{1+e^{-i\p{d-c}\c k}} \op{d}{c},
-\end{align}
-and define the vectors
+  &= \sum_{d\ne0} \sp{-2+2\p{N-1}} \op{d}
+  - \sum_{\substack{c,d\ne0\\c\ne d}} \p{1+e^{-i\p{d-c}\c k}} \op{d}{c} \\
+  &= 2\p{N-1} \1_{N-1}
+  - \sum_{c,d\ne0} \p{1+e^{-i\p{d-c}\c k}} \op{d}{c},
+\end{align}
+where $\1_{N-1}\equiv\sum_{d\ne0}\op{d}$ is the $\p{N-1}\times\p{N-1}$
+identity matrix, and define the vectors
 \begin{align}
@@
   \ut{\check{\m h}}_k
-  = -\p{N-1} \p{\op{\ut{0}} + \op{\ut{k}}}.
+  = 2\p{N-1} \1_{N-1} -\p{N-1} \p{\op{\ut{0}} + \op{\ut{k}}}.
 \end{align}
@@
   \ut{\check{\m h}}_k
-  = - \sum_{s\in\set{\pm}} \varepsilon_s \op{\ut{k}^s}
+  = 2\p{N-1} \1_{N-1} - \sum_{s\in\set{\pm}} \varepsilon_s \op{\ut{k}^s}
 \end{align}
@@
 \begin{align}
-  \ut{\check{\m h}}_0 = -2\p{N-1} \op{\ut{0}}.
+  \ut{\check{\m h}}_0 = 2\p{N-1} \p{\1_{N-1} - \op{\ut{0}}}.
 \end{align}
@@
 \begin{align}
-  \Delta_0 \equiv -2\p{N-1} - 2 h_0 = 0,
-  &&
-  \Delta_+ \equiv -\varepsilon_+ - 2 h_0 = N,
+  \Delta_0 \equiv 0,
+  &&
+  \Delta_+ \equiv 2\p{N-1} - \varepsilon_+ = N,
 \end{align}
@@
 \begin{align}
-  \Delta_\perp \equiv - 2 h_0 = 2N - 2.
+  \Delta_\perp \equiv 2\p{N-1}.
 \end{align}

### 2020-01-28 collective_sun/sun_perturbation.tex
fixed minor error in multi-body eigenvalue problem
simplifications to derivations of M-body eigenvalue problem and the two-body product

@@
   = E_0 \V_1 \ket\psi
-  + \sum_X \p{\sum_{p,q} h_{pq} v_{Xp} X_q - \sum_p h_p v_{Xp} X_p}
+  + \sum_X \sum_p \p{\sum_q h_{pq} v_{Xq} - h_p v_{Xp}} X_p
   \ket\psi,
@@
 \section{Diagnosing perturbed states with the interaction Hamiltonian}
-
-Here we simplify vectors of the form $H_0\V_M\ket\psi$ to arrive at
-the expansions in \eqref{eq:H_V1_psi} and \eqref{eq:H_V2_psi}.
+\label{sec:H_VM_psi}
+
+In this section, we simplify the products of the form
+$H_0\V_M\ket\psi$ to arrive at the expansions in \eqref{eq:H_V1_psi}
+and \eqref{eq:H_V2_psi}.  Here $H_0$ is the SU($n$)-symmetric
+interaction Hamiltonian in \eqref{eq:ints}, $\ket\psi$ is a fully
+symmetric state, and $\V_M$ is an $M$-body perturbation of the form
+\begin{align}
+  \V_M \equiv \sum_{k\in\C_N\p{M}} w_k O_k.
+\end{align}
+where $O$ is an $M$-body operator that is invariant under arbitrary
+permutation of its tensor factors; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is
+a choice of $M$ distinct spins that $O$ acts on; and $\C_N\p{n}$ is a
+choice of $n$ elements from a set of $N$.
 
@@
 
-We wish to simplify
+We first simplify
 \begin{align}
@@
 \end{align}
-where $\ket\psi\in\M_\FS$ is an arbitrary fully symmetric state, and
-without loss of generality we consider a single-body perturbation
-$\V_1$ built out of only one single-body operator $X$.  The sum in
-\eqref{eq:H_V1_psi_start} has terms with $k\in\set{p,q}$, and terms
-with $k\notin\set{p,q}$.  In the case of $k\notin\set{p,q}$, the
-permutation operator $\Pi_{pq}$ commutes with $X_k$ and annihilates on
-$\ket\psi$, leaving terms at each fixed $k$ of the form
+where without loss of generality we consider a single-body
+perturbation $\V_1$ built out of only one single-body operator $X$.
+The sum in \eqref{eq:H_V1_psi_start} has terms with $k\in\set{p,q}$,
+and terms with $k\notin\set{p,q}$.  In the case of $k\notin\set{p,q}$,
+the permutation operator $\Pi_{pq}$ commutes with $X_k$ and
+annihilates on $\ket\psi$, leaving terms at each fixed $k$ of the form
 \begin{align}
@@
   = E_0 \V_1 \ket\psi
-  + \p{\sum_{p,q} h_{pq} v_p X_q - \sum_p h_p v_p X_p}
-  \ket\psi.
+  + \sum_p \p{\sum_{q} h_{pq} v_q - h_p v_p}
+  X_p \ket\psi.
 \end{align}
@@
 
-Given an $M$-body perturbation
-\begin{align}
-  \V_M \equiv \sum_{k\in\C_N\p{M}} w_k O_k,
-\end{align}
-where $O$ is an $M$-body operator; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is
-a choice of $M$ distinct spins that $O$ acts on; and $\C_N\p{n}$ is a
-choice of $n$ elements from a set of $N$; we wish to simplify
+We wish to simplify
 \begin{align}
@@
 \end{align}
-where $\ket\psi\in\M_\FS$ is an arbitrary fully symmetric state.  The
-sum in \eqref{eq:H_VM_psi_start} has terms with $p,q\in k$, terms with
+where without loss of generality we consider an $M$-body perturbation
+$\V_M$ built out of only one $M$-body operator $O$.  The sum in
+\eqref{eq:H_VM_psi_start} has terms with $p,q\in k$, terms with
 $p,q\notin k$, and terms with only one of $p$ or $q\in k$.  The
@@
 $q$ with the restriction that $q\notin k$ into a sum over all $q$
-minus the remainder $q\in k$.  Defining $h_p\equiv\sum_q h_{pq}$, we thus find that
+minus the remainder $q\in k$.  We thus find that
 \begin{align}
   \sum_{\substack{\p{p,q}\in\C_N\p{2}\\p,q\in k~\t{or}~p,q\notin k}} h_{pq}
-  = E_0 + 2\sum_{\p{p,q}\subset k} h_{pq} - \sum_{p\in k} h_p,
+  = E_0 + \bar h_k,
+  &&
+  \bar h_k \equiv 2\sum_{\p{p,q}\subset k} h_{pq} - \sum_{p\in k} h_p,
+  &&
+  h_p \equiv \sum_q h_{pq},
 \end{align}
@@
   \p{h\circ w}_k
-  \equiv \sum_{j=1}^M \sum_{p\in k} h_{pk_j} w_{k_{jp}}
-  = h_{pk_1} w_{pk_2k_3\cdots k_M}
+  \equiv \sum_{\substack{p\in\ZZ_N\\p\notin k}} \sum_{j=1}^M
+  h_{pk_j} w_{k_{jp}}
+  = \sum_{\substack{p\in\ZZ_N\\p\notin k}}
+  \p{h_{pk_1} w_{pk_2k_3\cdots k_M}
     + h_{pk_2} w_{k_1pk_3\cdots k_M}
-  + h_{pk_M} \cdots w_{k_1k_2\cdots k_{M-1} p}
-\end{align}
-and
-\begin{align}
-  \bar h_k \equiv 2 \sum_{\p{p,q}\subset k} h_{pq} - \sum_{p\in k} h_p,
+    + \cdots + h_{pk_M} w_{k_1k_2\cdots k_{M-1} p}},
 \end{align}
@@
 \end{align}
+If $M=1$, then $\bar h_k=-h_k$ and
+$\p{h\circ w}_k=\sum_\ell h_{k\ell}w_\ell$, so
+\begin{align}
+  H_0 \V_1 \ket\psi
+  = E_0 \V_1 \ket\psi + \sum_{k\in\ZZ_N}
+  \p{\sum_\ell h_{k\ell} w_\ell - h_k w_k} O_k \ket\psi.
+\end{align}
+If $M=2$, then
+$\p{h\circ w}_{pq}=\sum_k\p{h_{kp} w_{kq}+h_{kq}w_{pk}}$, so
+\begin{align}
+  H_0 \V_2 \ket\psi
+  = E_0 \V_2 \ket\psi + \sum_{p<q}
+  \sp{\bar h_{pq} w_{pq} + \sum_k\p{h_{kp} w_{kq} + h_{kq} w_{pk}}}
+  O_k \ket\psi,
+\end{align}
+where $\bar h_{pq}=2h_{pq}-h_p-h_q$ with $h_p\equiv\sum_qh_{pq}$.
 
@@
   w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq}
-  &\EQFS \f14 \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}}
-  w_{Ok\ell} w_{Qkq} \p{O_{0,1} Q_{0,2} + O_{0,1} Q_{1,2}} \\
+  &\EQFS \f12 \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}}
+  w_{Ok\ell} w_{Qkq} O_{0,1} Q_{1,2} \\
   \f14 \sum_{\substack{k,\ell\\p\notin\set{k,\ell}\\q\in\set{k,\ell}}}
   w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq}
-  &\EQFS \f14 \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}}
-  w_{Ok\ell} w_{Qkq} \p{O_{0,1} Q_{2,0} + O_{0,1} Q_{2,1}},
+  &\EQFS \f12 \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}}
+  w_{Ok\ell} w_{Qkq} O_{0,1} Q_{1,2},
 \end{align}
@@
 \begin{align}
-  \f14 \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}} w_{Ok\ell} w_{Qkq}
-  = \f14 \sum_{k,\ell,q} w_{Ok\ell} w_{Qkq}
-  - \f14 \sum_{k,\ell} w_{Ok\ell} w_{Qk\ell}
-  = \f14 \v w_O\c\v w_Q - \f12 \v{\m w}_O \c \v{\m w}_Q.
+  \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}} w_{Ok\ell} w_{Qkq}
+  = \sum_{k,\ell,q} w_{Ok\ell} w_{Qkq}
+  - \sum_{k,\ell} w_{Ok\ell} w_{Qk\ell}
+  = \v w_O\c\v w_Q - 2 \v{\m w}_O \c \v{\m w}_Q.
 \end{align}
@@
   \P_0 \p{\V_2^\Delta}^2 \P_0
-  \EQFS - \sum_{O,Q} \bigg[
-  \p{\v w_O \c \v w_Q - \v{\m w}_O \c \v{\m w}_Q} O_{0,1} Q_{2,3}
-  - \v{\m w}_O \c \v{\m w}_Q O_{0,1} Q_{0,1} \\
-  -\f14 \p{\v w_O\c\v w_Q - 2 \v{\m w}_O \c \v{\m w}_Q}
-  O_{0,1} \p{Q_{0,2} + Q_{2,0} + Q_{1,2} + Q_{2,1}} \bigg].
+  \EQFS \sum_{O,Q}
+  \bigg[ \v{\m w}_O \c \v{\m w}_QO_{0,1} Q_{0,1}
+  + \p{\v w_O\c\v w_Q - 2 \v{\m w}_O \c \v{\m w}_Q}
+  O_{0,1} Q_{1,2} \\
+    - \p{\v w_O \c \v w_Q - \v{\m w}_O \c \v{\m w}_Q}
+    O_{0,1} Q_{2,3} \bigg].
   \label{eq:POQP}

### 2020-01-30 collective_sun/sun_perturbation.tex
added calculation of general product of multi-local operators restricted to the fully symmetric manifold

@@
 \renewcommand{\c}{\cdot} % inner product
-\renewcommand{\oc}{\circ} % element-wise product
 
@@
 
-\newcommand{\B}{\mathcal{B}}
 \newcommand{\C}{\mathcal{C}}
-\newcommand{\D}{\mathcal{D}}
 \newcommand{\E}{\mathcal{E}}
+\newcommand{\G}{\mathcal{G}}
 \renewcommand{\H}{\mathcal{H}}
 \newcommand{\I}{\mathcal{I}}
-\newcommand{\K}{\mathcal{K}}
 \renewcommand{\L}{\mathcal{L}}
 \newcommand{\M}{\mathcal{M}}
-\newcommand{\N}{\mathcal{N}}
 \renewcommand{\O}{\mathcal{O}}
 \renewcommand{\P}{\mathcal{P}}
-\newcommand{\Q}{\mathcal{Q}}
-\newcommand{\R}{\mathcal{R}}
-\newcommand{\T}{\mathcal{T}}
 \renewcommand{\S}{\mathcal{S}}
@@
 \newcommand{\X}{\mathcal{X}}
-\newcommand{\Y}{\mathcal{Y}}
-\newcommand{\Z}{\mathcal{Z}}
-
+
+\newcommand{\PP}{\mathbb{P}}
 \renewcommand{\SS}{\mathbb{S}}
 \newcommand{\TT}{\mathbb{T}}
-\newcommand{\VV}{\mathbb{V}}
 \newcommand{\ZZ}{\mathbb{Z}}
@@
 \setlist{nolistsep} % more compact spacing between environments
+
+% for drawing set (Venn) diagrams
+\usepackage{tikz}
+\tikzset{
+  baseline = (current bounding box.center)
+}
+\usetikzlibrary{math}
 
@@
   H_2^{(1)} = \sum_O \sum_{p<q} w_{Opq} \P_0 O_{pq} \P_0
-  \EQFS \sum_O \sum_{p<q} w_{Opq} O_{0,1}
+  \EQFS \sum_O \sum_{p<q} w_{Opq} O_{1,2}
   \EQFS \sum_O \bar w_O \col{O},
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Simplifying operator products restricted to the fully
-  symmetric manifold}
+\section{Operator products restricted to the fully symmetric manifold}
 \label{sec:sym_prod}
+
+Given a system of $N$ spins, we wish to simplify a product of
+multi-body operators restricted to the fully symmetric manifold.  For
+$p$ multi-body operators, such a product takes the form
+\begin{align}
+  \P_0 \sp{\prod_{j\in\ZZ_p}
+    \sum_{k\in\C_N\p{M_j}} w_j\p{k} O_j\p{k}} \P_0,
+  \label{eq:sym_prod_choice}
+\end{align}
+where $\P_0$ is a projector onto the fully symmetric manifold;
+$k\in\C_N\p{M_j}$ is a choice of $M_j$ distinct spins; $O_j\p{k}$
+denotes the action of a permutationally-symmetric $M_j$-body operator
+acting on spins $k$; and $w_j\p{k}$ is a scalar coefficient for
+$O_j\p{k}$.
+
+To simplify the product in \eqref{eq:sym_prod_choice}, we first write
+\begin{align}
+  \prod_{j\in\ZZ_p} \sum_{k\in\C_N\p{M_j}} w_j\p{k} O_j\p{k}
+  = \prod_{j\in\ZZ_p} \f1{M_j!} \sum_{k\in\ZZ_N^{M_j}} w_j\p{k} O_j\p{k},
+  \label{eq:sym_prod_vec}
+\end{align}
+where we extend the definitions of $w_j\p{k}$ and $O_j\p{k}$ to all
+vectors $k\in\ZZ_N^{M_j}$ by defining symmetric rank-$M_j$ tensors
+$w_j$ and $O_j$ (where $O_j$ is an {\it operator-valued} tensor) for
+which $w_j\p{k}=O_j\p{k}=0$ if $k$ contains any repeated symbols.
+
+In order to project \eqref{eq:sym_prod_vec} onto the fully symmetric
+manifold, we need to collect terms whose operator content is
+equivalent up to a permutation of spins.  To this end, we classify
+terms in \eqref{eq:sym_prod_vec} by the numbers $g_S$ of indices
+shared by all tensors $w_j,O_j$ with $j\in S\subset\ZZ_p$.  For
+example, a term of the form $w_1\p{a,b,c,d} w_2\p{b,c,d} w_3\p{b}$
+with distinct indices $\p{a,b,c,d}\in\C_N\p{4}$ would have
+\begin{align}
+  g_{\set{1}} = 1,
+  &&
+  g_{\set{1,2,3}} = 1,
+  &&
+  g_{\set{1,2}} = 2,
+\end{align}
+and $g_S=0$ for all other subsets $S\subset\ZZ_3$.  This index
+assignment can be represented by the Venn diagram
+\begin{align}
+  \begin{tikzpicture}]
+    \tikzmath{
+      \scale = 0.7;
+      \rr = 2.5 * \scale;
+      \aa = \scale;
+      \bb = sqrt(3) * \aa;
+      \cc = 2 * \aa;
+      \topY = \cc + \rr;
+      \crossLeftRightUpper = sqrt(\rr*\rr - \bb*\bb) - \aa;
+      \dotTopY = ( \crossLeftRightUpper + \topY ) / 2;
+      \dotTopLeftHeight = \crossLeftRightUpper - \scale/5;
+      \dotTopLeftXFst = -\rr*0.48 - \scale/3;
+      \dotTopLeftXSnd = -\rr*0.48 + \scale/3;
+      \labelDist = \scale * 0.8;
+      \labelTopY = \topY + \labelDist;
+      \labelBottomX = \labelTopY * sqrt(3) / 2;
+      \labelBottomY = \labelTopY / 2;
+    }
+    \draw (0, \cc em) circle [radius = \rr em];
+    \draw (-\bb em, -\aa em) circle [radius = \rr em];
+    \draw (+\bb em, -\aa em) circle [radius = \rr em];
+    \node[draw,fill,shape=circle,scale=0.3] at (0,0) {};
+    \node[draw,fill,shape=circle,scale=0.3] at (0, \dotTopY em) {};
+    \node[draw,fill,shape=circle,scale=0.3]
+    at (\dotTopLeftXFst em, \dotTopLeftHeight em) {};
+    \node[draw,fill,shape=circle,scale=0.3]
+    at (\dotTopLeftXSnd em, \dotTopLeftHeight em) {};
+    \node at (0, \labelTopY em) {1};
+    \node at (-\labelBottomX em, -\labelBottomY em) {2};
+    \node at (+\labelBottomX em, -\labelBottomY em) {3};
+  \end{tikzpicture},
+  \label{eq:venn_diagram}
+\end{align}
+where $g_S$ is determined the number of dots at the intersection of
+the circles $j\in S$.  We denote an assignment of distinct indices
+according to choices of $g_S$ for all $S\in\PP\p{\ZZ_p}$ by $g$, where
+$\PP\p{\ZZ_p}$ is the power set (i.e.~set of all subsets) of $\ZZ_p$.
+Given a vector $\m M$ of the tensor ranks $M_j$, we then denote the
+set of all valid index assignments $g$ by $\G\p{\m M}$.  For
+consistency, all valid index assignments $g\in\G\p{\m M}$ must assign
+exactly $M_j$ indices to the rank-$M_j$ tensor $w_j$, i.e.~
+\begin{align}
+  \sum_{S\in\PP\p{\ZZ_p}\,:\,j\in S} g_S
+  = \sum_{R\in\PP\p{\ZZ_p\setminus\set{j}}} g_{\set{j}\cup R}
+  = M_j
+\end{align}
+for all $j\in\ZZ_p$.  In order to make the index assignments
+determined by choices of $g_S$ for all $S\in\PP\p{\ZZ_p}$ unique, we
+also enforce that $g_{\set{}}=0$ for all $g\in\G\p{\m M}$, which is
+equivalent to saying that every dot in the corresponding Venn diagram
+must lie within some circle.  The set of valid index assignments
+$\G\p{\m M}$ is then essentially the set of all Venn diagrams of the
+form in \eqref{eq:venn_diagram} with $p=\rank\p{\m M}$ circles and a
+total of $M_j$ dots within circle $j$.
+
+A classification of terms in \eqref{eq:sym_prod_vec} by index
+assignments $g\in\G\p{\m M}$ allows us to expand
+\begin{align}
+  \prod_{j\in\ZZ_p} \sum_{k\in\ZZ_N^{M_j}} w_j\p{k} O_j\p{k}
+  \EQFS \sum_{g\in\G\p{\m M}} \S\p{g} w\p{g} O\p{g},
+  \label{eq:sym_prod_group_start}
+\end{align}
+where $O\p{g}$ is an operator acquired assigning indices to the
+tensors $O_j$ in a manner consistent with $g$; $w\p{g}$ is a scalar
+acquired by summing over all indices assigned to the tensors $w_j$
+according to $g$; and $\S\p{g}$ is a symmetry factor accounting for
+the number of equivalent ways to assign indices according to $g$.
+
+In order to write out the factors in \eqref{eq:sym_prod_group_start}
+explicitly, we define the set of values that the distinct indices
+assigned according to $g$ can take,
+\begin{align}
+  \C_N\p{g}
+  \equiv \set{ k \in \bigotimes_{S\in\PP\p{\ZZ_p}} \C_N\p{g_S}
+    : \t{all elements of $k$ are distinct} }
+  \simeq \C_N\p{\sum_{S\in\PP\p{\ZZ_p}}g_S},
+\end{align}
+where each $k\in\C_N\p{g}$ has the natural decomposition
+$k=\p{k_S:S\in\PP\p{\ZZ_p}}$ with $k_S\in\C_N\p{g_S}$, and we define
+the restriction of $k\in\C_N\p{g}$ to values associated with the
+tensors $w_j,O_j$,
+\begin{align}
+  k_j \equiv \bigcup_{S\in\PP\p{\ZZ_p} : j\in S} k_S
+  = \bigcup_{R\in\PP\p{\ZZ_p\setminus\set{j}}} k_{\set{j}\cup R},
+\end{align}
+where the union of lists $k_S$ denotes a concatenation of those lists.  We then have that
+\begin{align}
+  w\p{g} \equiv \sum_{k\in\C_N\p{g}} \prod_{j\in\ZZ_p} w_j\p{k_j},
+  &&
+  O\p{g}
+  \equiv \prod_{j\in\ZZ_p} O_j\p{k_j}~\t{for some}~k\in\C_N\p{g},
+\end{align}
+and
+\begin{align}
+  \S\p{g} \equiv \sp{\prod_{j\in\ZZ_p} { M_j \choose g_S : j\in S }}
+  \sp{\prod_{S\in\PP\p{\ZZ_p}} \p{g_S!}^\abs{S}}
+  = \prod_{j\in\ZZ_p} M_j!,
+  \label{eq:sym_prod_symmetry}
+\end{align}
+where the first product in \eqref{eq:sym_prod_symmetry} contains multinomial coefficients
+\begin{align}
+  { M \choose a_1, a_2, \cdots, a_m }
+  \equiv \f{M!}{a_1!a_2!\cdots a_m!},
+\end{align}
+that account for the number of ways to partition the indices of each
+$w_j,O_j$ into sets of shared indices according to $g_S$ for all
+$j\in S$, and the second product in \eqref{eq:sym_prod_symmetry}
+accounts for the number of ways to permute each set of shared indices
+on every tensor.  The choice of $k\in\C_N\p{g}$ in the definition of
+$O\p{g}$ is irrelevant after projection onto the fully symmetric
+manifold.  Altogether, we have that
+\begin{align}
+  \prod_{j\in\ZZ_p} \sum_{k\in\C_N\p{M_j}} w_j\p{k} O_j\p{k}
+  \EQFS \sum_{g\in\G\p{\m M}} w\p{g} O\p{g}.
+  \label{eq:sym_prod_group}
+\end{align}
 
@@
   w_{Ok\ell} w_{Qk\ell} O_{k\ell} Q_{k\ell}
-  = \v{\m w}_O \c \v{\m w}_Q O_{0,1} Q_{0,1}.
+  = \v{\m w}_O \c \v{\m w}_Q O_{1,2} Q_{1,2}.
 \end{align}
@@
   &\EQFS \f12 \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}}
-  w_{Ok\ell} w_{Qkq} O_{0,1} Q_{1,2} \\
+  w_{Ok\ell} w_{Qkq} O_{1,2} Q_{2,3} \\
   \f14 \sum_{\substack{k,\ell\\p\notin\set{k,\ell}\\q\in\set{k,\ell}}}
@@
   &\EQFS \f12 \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}}
-  w_{Ok\ell} w_{Qkq} O_{0,1} Q_{1,2},
+  w_{Ok\ell} w_{Qkq} O_{1,2} Q_{2,3},
 \end{align}
@@
   \EQFS \f14 \sum_{\substack{k,\ell\\p,q\notin\set{k,\ell}}}
-  w_{Ok\ell} w_{Qpq} O_{0,1} Q_{2,3},
+  w_{Ok\ell} w_{Qpq} O_{1,2} Q_{3,4},
 \end{align}
@@
   \EQFS \sum_{O,Q}
-  \bigg[ \v{\m w}_O \c \v{\m w}_QO_{0,1} Q_{0,1}
+  \bigg[ \v{\m w}_O \c \v{\m w}_Q O_{1,2} Q_{1,2}
   + \p{\v w_O\c\v w_Q - 2 \v{\m w}_O \c \v{\m w}_Q}
-  O_{0,1} Q_{1,2} \\
+  O_{1,2} Q_{2,3} \\
     - \p{\v w_O \c \v w_Q - \v{\m w}_O \c \v{\m w}_Q}
-    O_{0,1} Q_{2,3} \bigg].
+    O_{1,2} Q_{3,4} \bigg].
   \label{eq:POQP}
@@
 turns out to be rather non-trivial in general.  We therefore discuss
-the decomposition of multi-local operators (e.g.~$O_{0,1} Q_{1,2}$, a
+the decomposition of multi-local operators (e.g.~$O_{1,2} Q_{2,3}$, a
 3-local operator) into a sum of collective operators within the fully
@@
 sites are indexed by vectors
-$p\in\ZZ_L^D\simeq\set{0,1,\cdots,L-1}^D$, and wavenumbers take on
-values $k\in\ZZ_L^D\times2\pi/L$.  The corresponding eigenvalues of
-$\m h$ can be determined by expanding
+$p\in\ZZ_L^D\simeq\set{0,1,\cdots,L-1}^{\otimes D}$, and wavenumbers
+take on values $k\in\ZZ_L^D\times2\pi/L$.  The corresponding
+eigenvalues of $\m h$ can be determined by expanding
 \begin{align}
@@
 \begin{align}
-  \bar{\m h}*\m w + \P_{\t{off-diag}}\p{\m h\c\m w + \m w\c\m h}
+  \bar{\m h} * \m w + \P_{\t{off-diag}}\p{\m h\c\m w + \m w\c\m h}
   = \Delta \m w,

### 2020-01-30 collective_sun/sun_perturbation.tex
improed language in last derivation

@@
 
-Given a system of $N$ spins, we wish to simplify a product of
-multi-body operators restricted to the fully symmetric manifold.  For
-$p$ multi-body operators, such a product takes the form
+Given a system of $N$ spins, we wish to project a product of
+multi-body operators onto the fully symmetric manifold.  For $p$
+multi-body operators, the projection of such a product takes the form
 \begin{align}
@@
     \sum_{k\in\C_N\p{M_j}} w_j\p{k} O_j\p{k}} \P_0,
-  \label{eq:sym_prod_choice}
+  \label{eq:sym_prod_proj}
 \end{align}
@@
 $k\in\C_N\p{M_j}$ is a choice of $M_j$ distinct spins; $O_j\p{k}$
-denotes the action of a permutationally-symmetric $M_j$-body operator
-acting on spins $k$; and $w_j\p{k}$ is a scalar coefficient for
-$O_j\p{k}$.
-
-To simplify the product in \eqref{eq:sym_prod_choice}, we first write
+denotes the action of an $M_j$-body operator on spins $k$ that is
+invariant under permutations of $k$; and $w_j\p{k}$ is a scalar
+coefficient for $O_j\p{k}$.
+
+To simplify the product in \eqref{eq:sym_prod_proj}, we first write
 \begin{align}
   \prod_{j\in\ZZ_p} \sum_{k\in\C_N\p{M_j}} w_j\p{k} O_j\p{k}
-  = \prod_{j\in\ZZ_p} \f1{M_j!} \sum_{k\in\ZZ_N^{M_j}} w_j\p{k} O_j\p{k},
+  = \prod_{j\in\ZZ_p} \f1{M_j!} \sum_{k\in\ZZ_N^{M_j}}
+  w_j\p{k} O_j\p{k},
   \label{eq:sym_prod_vec}
@@
 $w_j$ and $O_j$ (where $O_j$ is an {\it operator-valued} tensor) for
-which $w_j\p{k}=O_j\p{k}=0$ if $k$ contains any repeated symbols.
-
-In order to project \eqref{eq:sym_prod_vec} onto the fully symmetric
-manifold, we need to collect terms whose operator content is
-equivalent up to a permutation of spins.  To this end, we classify
-terms in \eqref{eq:sym_prod_vec} by the numbers $g_S$ of indices
-shared by all tensors $w_j,O_j$ with $j\in S\subset\ZZ_p$.  For
-example, a term of the form $w_1\p{a,b,c,d} w_2\p{b,c,d} w_3\p{b}$
-with distinct indices $\p{a,b,c,d}\in\C_N\p{4}$ would have
+which $w_j\p{k}=O_j\p{k}=0$ if $k$ contains any repeated symbols.  We
+then collect terms whose operator content is equivalent up to a
+permutation of all spins.  To this end, we classify terms in
+\eqref{eq:sym_prod_vec} by the numbers $g_S$ of indices shared by all
+tensors $w_j,O_j$ with $j\in S\subset\ZZ_p$.  For example, a term of
+the form $w_1\p{a,b,c,d} w_2\p{b,c,d} w_3\p{b}$ with distinct indices
+$\p{a,b,c,d}\in\C_N\p{4}$ would have
 \begin{align}
@@
 \end{align}
-where $O\p{g}$ is an operator acquired assigning indices to the
+where $O\p{g}$ is an operator acquired by assigning indices to the
 tensors $O_j$ in a manner consistent with $g$; $w\p{g}$ is a scalar
@@
 In order to write out the factors in \eqref{eq:sym_prod_group_start}
-explicitly, we define the set of values that the distinct indices
-assigned according to $g$ can take,
+explicitly, we define
+\begin{align}
+  \abs{g} \equiv \sum_{S\in\PP\p{\ZZ_p}} g_S,
+\end{align}
+which is equivalent to the total number of dots in the Venn diagram
+associated with $g$, and identify the set of values that the distinct
+indices assigned according to $g$ can take:
 \begin{align}
@@
     : \t{all elements of $k$ are distinct} }
-  \simeq \C_N\p{\sum_{S\in\PP\p{\ZZ_p}}g_S},
+  \simeq \C_N\p{\abs{g}},
 \end{align}
 where each $k\in\C_N\p{g}$ has the natural decomposition
-$k=\p{k_S:S\in\PP\p{\ZZ_p}}$ with $k_S\in\C_N\p{g_S}$, and we define
-the restriction of $k\in\C_N\p{g}$ to values associated with the
-tensors $w_j,O_j$,
+$k=\p{k_S:S\in\PP\p{\ZZ_p}}$ with $k_S\in\C_N\p{g_S}$.  We then define
+the restriction of $k\in\C_N\p{g}$ to values associated with $w_j$ and
+$O_j$:
 \begin{align}
@@
 \end{align}
-where the union of lists $k_S$ denotes a concatenation of those lists.  We then have that
+where the union of lists $k_S$ denotes a concatenation of those lists.
+We then have that
 \begin{align}
@@
   &&
-  O\p{g}
-  \equiv \prod_{j\in\ZZ_p} O_j\p{k_j}~\t{for some}~k\in\C_N\p{g},
-\end{align}
-and
+  O\p{g} \equiv \prod_{j\in\ZZ_p}
+  O_j\p{\ell_j}~\t{for some}~\ell\in\C_N\p{g},
+\end{align}
+where the choice of $\ell\in\C_N\p{g}$ does not matter after
+projection onto fully symmetric manifold, and
 \begin{align}
@@
 \end{align}
-where the first product in \eqref{eq:sym_prod_symmetry} contains multinomial coefficients
+where the first product in \eqref{eq:sym_prod_symmetry} contains
+multinomial coefficients
 \begin{align}
@@
 accounts for the number of ways to permute each set of shared indices
-on every tensor.  The choice of $k\in\C_N\p{g}$ in the definition of
-$O\p{g}$ is irrelevant after projection onto the fully symmetric
-manifold.  Altogether, we have that
+on every tensor.  Altogether, the symmetry factors $\S\p{g}$ turn out
+to be independent of $g$, and cancel out exactly with the $\sim M_j!$
+prefactors in \eqref{eq:sym_prod_vec}, so
 \begin{align}

### 2020-01-30 collective_sun/sun_perturbation.tex
fix mistake in last derivation

@@
 In order to write out the factors in \eqref{eq:sym_prod_group_start}
-explicitly, we define
-\begin{align}
-  \abs{g} \equiv \sum_{S\in\PP\p{\ZZ_p}} g_S,
-\end{align}
-which is equivalent to the total number of dots in the Venn diagram
-associated with $g$, and identify the set of values that the distinct
-indices assigned according to $g$ can take:
+explicitly, we identify the set of values that the distinct indices
+assigned according to $g$ can take:
 \begin{align}
@@
   \equiv \set{ k \in \bigotimes_{S\in\PP\p{\ZZ_p}} \C_N\p{g_S}
-    : \t{all elements of $k$ are distinct} }
-  \simeq \C_N\p{\abs{g}},
+    : \t{all elements of $k$ are distinct} },
 \end{align}
@@
 where the union of lists $k_S$ denotes a concatenation of those lists.
-We then have that
+These definitions allow us to expand
 \begin{align}
@@
 accounts for the number of ways to permute each set of shared indices
-on every tensor.  Altogether, the symmetry factors $\S\p{g}$ turn out
-to be independent of $g$, and cancel out exactly with the $\sim M_j!$
-prefactors in \eqref{eq:sym_prod_vec}, so
+on every tensor.  Altogether, the symmetry factors $\S\p{g}$ cancel
+out with the $\sim M_j!$ prefactors in \eqref{eq:sym_prod_vec}, so
 \begin{align}

### 2020-01-30 collective_sun/sun_perturbation.tex
modified some derivations in the appendix on operator products

@@
   H_1^{(1)} = \sum_{X,p} v_{Xp} \P_0 X_p \P_0
-  \EQFS \sum_{X,p} v_{Xp} X_0
+  \EQFS \sum_{X,p} v_{Xp} X_1
   \EQFS \sum_X \bar v_X \col{X},
@@
 \end{align}
-where $O\p{g}$ is an operator acquired by assigning indices to the
-tensors $O_j$ in a manner consistent with $g$; $w\p{g}$ is a scalar
-acquired by summing over all indices assigned to the tensors $w_j$
-according to $g$; and $\S\p{g}$ is a symmetry factor accounting for
-the number of equivalent ways to assign indices according to $g$.
+where $\EQFS$ denotes equality up to a restriction to the fully
+symmetric manifold, $O\p{g}$ is an operator acquired by assigning
+indices to the tensors $O_j$ in a manner consistent with $g$; $w\p{g}$
+is a scalar acquired by summing over all indices assigned to the
+tensors $w_j$ according to $g$; and $\S\p{g}$ is a symmetry factor
+accounting for the number of equivalent ways to assign indices
+according to $g$.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Single-body product}
+\subsection{Product of two single-body operators}
 \label{sec:PXYP}
@@
 on the fully symmetric manifold $\M_0$ by the perturbation $\V_1$ in
-\eqref{eq:perturbations} requires us to simplify
-\begin{align}
-  \P_0 \p{\V_1^\Delta}^2 \P_0
-  \EQFS \sum_{\substack{X,Y\\p,q}} v_{Xp} v_{Yq} X_p Y_q
-  = \sum_{X,Y,p} v_{Xp} v_{Yp} X_p Y_p
-  + \sum_{\substack{X,Y\\p\ne q}} v_{Xp} v_{Yq} X_p Y_q,
+\eqref{eq:perturbations} requires us to simplify a product of the form
+\begin{align}
+  \P_0 \sp{\sum_{p,q} v_{Xp} v_{Yq} X_p Y_q} \P_0
+  \EQFS \sum_p v_{Xp} v_{Yp} X_1 Y_1
+  + \sum_{p\ne q} v_{Xp} v_{Yq} X_1 Y_2,
   \label{eq:PXYP_start}
 \end{align}
-where $\EQFS$ denotes equality under a restriction to the fully
-symmetric manifold $\M_0$; and all coefficient vectors $\v v_X$
-satisfy \eqref{eq:cond_1} with eigenvalue $\Delta\ne0$, which implies
-that all $\v v_X$ are mean-zero.  The first sum on the right of
-\eqref{eq:PXYP_start} is then
-\begin{align}
-  \sum_{X,Y,p} v_{Xp} v_{Yp} X_p Y_p
-  \EQFS \sum_{X,Y,p} v_{Xp} v_{Yp} X_0 Y_0
-  \EQFS \f1N \sum_{X,Y} \v v_X \c\v v_Y \col{XY},
+where $X,Y$ are single-body operators; and the coefficients
+$v_{Xp},v_{Yq}$ satisfy \eqref{eq:cond_1} with eigenvalue
+$\Delta\ne0$, which implies that they are mean-zero.  The first sum on
+the right of \eqref{eq:PXYP_start} is then
+\begin{align}
+  \sum_p v_{Xp} v_{Yp} X_1 Y_1
+  \EQFS \f1N \v v_X \c \v v_Y\, \col{XY},
   &&
   \col{Z} \equiv \sum_p Z_p.
-  \label{eq:PXYP_eq}
-\end{align}
-The second sum on the right of \eqref{eq:PXYP_start}, meanwhile, is
-\begin{align}
-  \sum_{\substack{X,Y\\p\ne q}} v_{Xp} v_{Yq} X_p Y_q
-  \EQFS \sum_{\substack{X,Y\\p\ne q}} v_{Xp} v_{Yq} X_0 Y_1,
-  \label{eq:PXYP_neq_start}
-\end{align}
-where we can use the fact that the vectors $\v v_X,\v v_Y$ are
-mean-zero to simplify the sum over $p\ne q$ in
-\eqref{eq:PXYP_neq_start} to
+\end{align}
+To simplify the second sum on the right of \eqref{eq:PXYP_start}, we
+use the fact that $\v v$ is mean-zero to write
 \begin{align}
@@
   = \sum_{p,q} v_{Xp} v_{Yq} - \sum_{p=q} v_{Xp} v_{Yq}
-  = - \v v_X\c\v v_Y,
-\end{align}
-In order to simplify the product of operators in
-\eqref{eq:PXYP_neq_start}, we expand
-\begin{align}
-  \col{X}\,\col{Y} = \sum_{p,q} X_p Y_q
+  = - \v v_X \c\v v_Y,
+\end{align}
+and we expand
+\begin{align}
+  \col{X}\,\col{Y}
   = \sum_{p=q} X_p Y_q + \sum_{p\ne q} X_p Y_q
-  \EQFS \col{XY} + N\p{N-1} X_0 Y_1,
-  \label{eq:PXYP_neq_ops}
+  \EQFS \col{XY} + N\p{N-1} X_1 Y_2,
 \end{align}
@@
 \begin{align}
-  \sum_{\substack{X,Y\\p\ne q}} v_{Xp} v_{Yq} X_p Y_q
-  \EQFS - \sum_{X\ne Y} \v v_X\c\v v_Y X_0 Y_1
-  \EQFS - \f1{N\p{N-1}}\sum_{X,Y} \v v_X\c\v v_Y
-  \p{\col{X}\,\col{Y} - \col{XY}}.
-  \label{eq:PXYP_neq}
+  \sum_{p\ne q} v_{Xp} v_{Yq} X_1 Y_2
+  \EQFS - \v v_X\c\v v_Y \p{\col{X}\,\col{Y} - \col{XY}}.
 \end{align}
@@
 \begin{align}
-  \P_0 \p{\V_1^\Delta}^2 \P_0
-  = - \f1{N\p{N-1}} \sum_{X,Y} \v v_X \c\v v_Y
+  \sum_{p,q} v_{Xp} v_{Yq} X_p Y_q
+  \EQFS - \f{\v v_X \c\v v_Y}{N\p{N-1}}
   \p{\col{X}\,\col{Y} - N \col{XY}}.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Two-body product}
+\subsection{Product of two two-body operators}
 \label{sec:POQP}
@@
 \eqref{eq:perturbations} requires us to simplify
-\begin{align}
-  \P_0 \p{\V_2^\Delta}^2 \P_0
-  = \sum_{O,Q} \sum_{\substack{k<\ell\\p<q}}
-  w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq}
-  \EQFS \f14 \sum_{O,Q} \sum_{k,\ell,p,q}
-  w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq},
+\begin{multline}
+  \P_0 \sp{\sum_{\substack{k<l\\p<q}}
+    w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq}} \P_0
+  \EQFS \sum_{p<q} w_{Opq} w_{Qpq} O_{1,2} Q_{1,2}
+  + \sum_{\substack{k,p,q\\k\ne p\ne q\ne k}}
+  w_{Okp} w_{Qkq} O_{1,2} Q_{1,3} \\
+  + \sum_{\substack{k<\ell\\p<q\\\set{k,\ell}\ne\set{p,q}}}
+  w_{Ok\ell} w_{Qpq} O_{1,2} Q_{3,4},
   \label{eq:POQP_start}
-\end{align}
-where $\EQFS$ denotes equality under a restriction to the fully
-symmetric manifold $\M_0$; we define $w_{Opq}$ for all $p,q$ (as
-opposed to only $p<q$) by enforcing $w_{Opq}=w_{Oqp}$ and $w_{Opp}=0$;
-and the coefficient vectors $\v w_O$ satisfy the eigenvalue equation
-\eqref{eq:cond_2} with eigenvalue $\Delta\ne0$, which implies that all
-$\v{\m w}_O$ are mean-zero.  To simplify the sum in
-\eqref{eq:POQP_start}, we first consider the terms with
-$\p{p,q}=\p{k,\ell}$:
-\begin{align}
-  \sum_{k<\ell}
-  w_{Ok\ell} w_{Qk\ell} O_{k\ell} Q_{k\ell}
-  = \v{\m w}_O \c \v{\m w}_Q O_{1,2} Q_{1,2}.
-\end{align}
-We then consider terms $p\in\set{k,\ell}$ and $q\notin\set{k,\ell}$ or
-vice versa:
-\begin{align}
-  \f14 \sum_{\substack{k,\ell\\p\in\set{k,\ell}\\q\notin\set{k,\ell}}}
-  w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq}
-  &\EQFS \f12 \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}}
-  w_{Ok\ell} w_{Qkq} O_{1,2} Q_{2,3} \\
-  \f14 \sum_{\substack{k,\ell\\p\notin\set{k,\ell}\\q\in\set{k,\ell}}}
-  w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq}
-  &\EQFS \f12 \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}}
-  w_{Ok\ell} w_{Qkq} O_{1,2} Q_{2,3},
-\end{align}
-where
-\begin{align}
-  \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}} w_{Ok\ell} w_{Qkq}
-  = \sum_{k,\ell,q} w_{Ok\ell} w_{Qkq}
-  - \sum_{k,\ell} w_{Ok\ell} w_{Qk\ell}
-  = \v w_O\c\v w_Q - 2 \v{\m w}_O \c \v{\m w}_Q.
-\end{align}
-Finally, the terms in \eqref{eq:POQP_start} with
-$p,q\notin\set{k,\ell}$ are
-\begin{align}
-  \f14 \sum_{\substack{k,\ell\\p,q\notin\set{k,\ell}}}
-  w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq}
-  \EQFS \f14 \sum_{\substack{k,\ell\\p,q\notin\set{k,\ell}}}
-  w_{Ok\ell} w_{Qpq} O_{1,2} Q_{3,4},
-\end{align}
-where, using the fact that all $\v{\m w}_O$ are mean-zero,
-\begin{multline}
-  \f14 \sum_{\substack{k,\ell\\p,q\notin\set{k,\ell}}} w_{Ok\ell} w_{Qpq}
-  = - \f14 \sum_{\substack{k,\ell\\p,q\in\set{k,\ell}}} w_{Ok\ell} w_{Qpq}
-  - \f14 \sum_{\substack{k,\ell\\p\in\set{k,\ell}\\q\notin\set{k,\ell}}}
-  w_{Ok\ell} w_{Qpq}
-  - \f14 \sum_{\substack{k,\ell\\p\notin\set{k,\ell}\\q\in\set{k,\ell}}}
-  w_{Ok\ell} w_{Qpq} \\
-  = -\v w_O \c \v w_Q + \v{\m w}_O \c \v{\m w}_Q.
 \end{multline}
+where $O$ and $Q$ are permutationally-symmetric two-body operators;
+and the coefficients $w_{Ok\ell},w_{Qpq}$ satisfy \eqref{eq:cond_2}
+with eigenvalue $\Delta\ne0$, which implies that they are mean-zero.
+Defining
+\begin{align}
+  \v{\m m} \equiv \sum_{p<q} m_{pq} \ket{\set{p,q}},
+  &&
+  \v m \equiv \sum_p m_p \ket{p},
+  &&
+  m_p \equiv \sum_q m_{pq},
+\end{align}
+we can write the first sum on the right of \eqref{eq:POQP_start} as
+\begin{align}
+  \sum_{p<q} w_{Opq} w_{Qpq} = \v{\m w}_O \c\v{\m w}_Q,
+\end{align}
+and, using the fact that $w_{Opp}=0$, we can write the second sum on
+the right of \eqref{eq:POQP_start} as
+\begin{align}
+  \sum_{\substack{k,p,q\\k\ne p\ne q\ne k}} w_{Okp} w_{Qkq}
+  = \sum_{k,p} w_{Okp} \sum_{q\ne p} w_{Okq}
+  = \sum_{k,p} w_{Okp} \p{w_{Qk} - w_{Qkp}}
+  = \v w_O \c \v w_Q - 2 \v{\m w}_O \c \v{\m w}_Q.
+\end{align}
+Using additionally the fact that $w_{Opq},w_{Qpq}$ are mean-zero, we
+can similarly simplify
+\begin{align}
+  \sum_{\substack{p<q\\k<\ell\\\set{k,\ell}\ne\set{p,q}}}
+  w_{Ok\ell} w_{Qpq} = -\v w_O \c \v w_Q + \v{\m w}_O \c \v{\m w}_Q.
+\end{align}
 Altogether, we thus have that
 \begin{multline}
-  \P_0 \p{\V_2^\Delta}^2 \P_0
-  \EQFS \sum_{O,Q}
-  \bigg[ \v{\m w}_O \c \v{\m w}_Q O_{1,2} Q_{1,2}
-  + \p{\v w_O\c\v w_Q - 2 \v{\m w}_O \c \v{\m w}_Q}
-  O_{1,2} Q_{2,3} \\
-    - \p{\v w_O \c \v w_Q - \v{\m w}_O \c \v{\m w}_Q}
-    O_{1,2} Q_{3,4} \bigg].
+  \sum_{\substack{k<l\\p<q}} w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq}
+  \EQFS \v{\m w}_O \c \v{\m w}_Q O_{1,2} Q_{1,2}
+  + \p{\v w_O\c\v w_Q - 2 \v{\m w}_O \c \v{\m w}_Q} O_{1,2} Q_{1,3} \\
+  - \p{\v w_O \c \v w_Q - \v{\m w}_O \c \v{\m w}_Q} O_{1,2} Q_{3,4}.
   \label{eq:POQP}

### 2020-01-30 collective_sun/sun_perturbation.tex
commented out obsolete section on converting between multi-local and collective operators

@@
 where the product $\P_0 \p{\V_2^\Delta}^2 \P_0$ is simplified in
-Appendices \ref{sec:POQP} and \ref{sec:multi_to_collective}.  In the
-special case that
+Appendix \ref{sec:POQP}.  In the special case that
 \begin{align}
@@
 \end{multline}
-In principle, our next task is to write this result in terms of
-collective operators within the fully symmetric manifold; this task
-turns out to be rather non-trivial in general.  We therefore discuss
-the decomposition of multi-local operators (e.g.~$O_{1,2} Q_{2,3}$, a
-3-local operator) into a sum of collective operators within the fully
-symmetric manifold in Appendix \ref{sec:multi_to_collective}.  At the
-end of Appendix \ref{sec:multi_to_collective}, we provide analytical
-expressions for the special cases of
-\begin{enumerate*}
-\item multi-local operators that are tensor powers of a single
-  operator, i.e.~of the form $X^{\otimes k}$; and
-\item multi-local operators of the form $X^{\otimes k}$ with $X^2=1$.
-\end{enumerate*}
-These cases are relevant for e.g.~the SU(2) case with Ising-type
-interactions of the form $Z\otimes Z$, with $Z$ a single-spin Pauli
-operator.
 
@@
 coefficient vector $\v{\m w}=\sum_\Delta\v{\m w}_\Delta$ into
-coefficient vectors $\v{\m w}_\Delta$ that generate a fixed-energy
-perturbation $\V_2^\Delta$ is then performed by projecting each
-$k$-component of $\v{\m w}$ onto the corresponding $\Delta$-eigenspace
-of $\ut{\check{\m h}}_k$.  If the perturbation $\V_2$ is
-translationally invariant, then the $k$-components of $\v{\m w}$ for
-all $k\ne 0$ all vanish, so we only need to consider the case of
-$k=0$.
+coefficient vectors $\v{\m w}_\Delta$ that generate a
+fixed-excitation-energy perturbation $\V_2^\Delta$ is then performed
+by projecting each $k$-component of $\v{\m w}$ onto the corresponding
+$\Delta$-eigenspace of $\ut{\check{\m h}}_k$.  If the perturbation
+$\V_2$ is translationally invariant, then the $k$-components of
+$\v{\m w}$ for all $k\ne 0$ all vanish, so we only need to consider
+the case of $k=0$.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Collective and multi-local operators in the fully symmetric
-  manifold}
-\label{sec:multi_to_collective}
-
-Here we provide a decomposition multi-local operators into sums of
-collective operators (and vice versa) within the fully symmetric
-manifold of $N$ spins.  Multi-local operators are operators of the
-form
-\begin{align}
-  O_S \equiv \bigotimes_{O\in S} O,
-  \label{eq:multi_op_set}
-\end{align}
-where $S\equiv\set{O}$ is a set of single-spin operators $O$.  The
-order of tensor factors in \eqref{eq:multi_op_set} does not matter
-after projecting operators onto the fully symmetric manifold, as we
-always will in this section.  While the decomposition of a collective
-operator into a sum of multi-local operators is unique, the opposite
-is not true: there is generally no unique decomposition of a
-multi-local operator into a sum of collective operators.  We can,
-however, enforce a canonical decomposition by choosing a set of
-collective operators that obey the same symmetries as $O_S$, namely
-invariance under arbitrary permutations of the single-spin operators
-$O\in S$.
-
-Even though $S$ contains $\abs{S}$ spins, a collective operator built
-from the single-spin operators $O\in S$ may have multiple
-(single-spin) operators acting on the same spin, resulting in a
-$k$-local operator with $k<\abs{S}$.  We therefore {\it
-  partition}\footnote{A partition $P\equiv\set{p}$ of $S$ is a set of
-  mutually disjoint, non-empty subsets $p\subset S$ with
-  $\bigcup_{p\in P}p=S$.} $S$ into subsets of operators that act on
-the same spin.  For any partition $P\equiv\set{p}$ of $S$ (here
-$p\subset S$ is a subset of operators $O\in S$ that address the same
-spin), we then define the $\abs{P}$-local operator
-\begin{align}
-  O_P \equiv \bigotimes_{p\in P} \bigodot_{O\in p} O,
-\end{align}
-where $\bigodot_{x\in\X} x$ denotes a symmetrized product of all
-elements in $\X$, i.e.
-\begin{align}
-  \bigodot_{x\in\X} x
-  \equiv \f1{\abs{\X}!}
-  \sum_{\substack{\t{permutations}\\\sigma~\t{of}~\X}}
-  ~ \prod_{k=1}^{\abs{\X}} x_{\sigma_k},
-\end{align}
-with $\sigma_k$ the $k$-th element of the permutation $\sigma$ when
-expressed in one-line notation; for example,
-\begin{align}
-  \bigodot_{j=1}^2 x_j &= x_1 \odot x_2 = \f12 \p{x_1x_2 + x_2x_1}, \\
-  \bigodot_{j=1}^3 x_j
-  &= \f1{3!} \p{x_1x_2x_3 + x_1x_3x_2 + x_2x_1x_3
-    + x_2x_3x_1 + x_3x_1x_2 + x_3x_2x_1}.
-\end{align}
-The set of multi-local operators $\set{O_P:\t{partitions}~P~\t{of}~S}$
-consists of all operators that
-\begin{enumerate*}
-\item can be built from ordinary products and tensor products of all
-  single-spin operators $O\in S$, and
-\item are invariant under arbitrary permutations of $S$.
-\end{enumerate*}
-For any set $S$ of single-spin operators and any partition $P$ of $S$,
-we can similarly define the collective operators
-\begin{align}
-  \O_S \equiv \bigodot_{O\in S} \col{O},
-  &&
-  \O_P \equiv \bigodot_{p\in P} \col{\textstyle\bigodot_{O\in p} O},
-\end{align}
-where we underline a single-spin operator to denote its collective
-version, i.e.~$\col{Q}\equiv\sum_j Q^{(j)}$, where $Q^{(j)}$ is an
-operator that acts with the single-spin operator $Q$ on spin $j$ and
-trivially on all other spins.  By construction, the set of collective
-operators $\set{\O_P:\t{partitions}~P~\t{of}~S}$ is one-to-one to the
-set of multi-local operators $\set{O_P:\t{partitions}~P~\t{of}~S}$,
-and it is straightforward to work out that for any set $S$ of
-single-spin operators,
-\begin{align}
-  \O_S
-  \EQFS \sum_{\substack{\t{partitions}\\P~\t{of}~S}}
-  N!_{\abs{P}} O_P,
-  &&
-  N!_k \equiv \prod_{j=0}^{k-1}\p{N-j}! = \f{N!}{\p{N-k}!},
-  \label{eq:collective_to_multi_full}
-\end{align}
-where $\EQFS$ denotes equality under a restriction to the fully
-symmetric manifold.  For any multi-local operator $O_S$ indexed by a
-set $S$ of single-spin operators, we can collect expansions of the
-form \eqref{eq:collective_to_multi_full} for all collective operators
-built from non-empty subsets of $S$ into a system of linear equations
-$\L$, and in turn solve $\L$ for $O_S$.  When
-$\abs{S}\in\set{1,2,3,4}$ (and, by conjecture, for all $S$), by use of
-a computer algebra system (namely, Mathematica) we find that
-\begin{align}
-  N!_{\abs{S}} O_S
-  &\EQFS \sum_{\substack{\t{partitions}\\P~\t{of}~S}}
-  \p{-1}^{\abs{S}-\abs{P}} \O_P \times \prod_{p\in P}\p{\abs{p}-1}!
-  \label{eq:multi_to_collective_full} \\
-  &\EQFS \f1{\abs{S}!} \sum_{\substack{\t{permutations}\\\Pi~\t{of}~S}}
-  \sum_{\substack{\t{ordered}\\\t{partitions}\\\tilde P~\t{of}~\Pi}}
-  \p{-1}^{\abs{S}-\abs{\tilde P}} \O_{\tilde P}
-  \times \prod_{\tilde p\in\tilde P} \p{\abs{\tilde p}-1}!
-  \label{eq:multi_to_collective_full_simp}
-\end{align}
-where we identify permutations $\Pi$ of $S$ with an ordered list
-(tuple) $\p{O_1,O_2,\cdots,O_{\abs{S}}}$ of all single-spin operators
-in $S$; and we enforce that partitions $\tilde P$ of an ordered list
-$\Pi$, as well as parts $\tilde p\in\tilde P$, are lexicographically
-ordered in accordance with the ordering of operators in $\Pi$.  If
-$S=\set{X,Y,Z}$, for example, then the ordered partitions of the
-trivial permutation $\Pi=\p{X,Y,Z}$ are
-\begin{align}
-  \p{\p{X,Y,Z}}, &&
-  \p{\p{X,Y},\p{Z}}, &&
-  \p{\p{X,Z},\p{Y}}, &&
-  \p{\p{X},\p{Y,Z}}, &&
-  \p{\p{X},\p{Y},\p{Z}}.
-\end{align}
-For ordered lists $\tilde S$ and ordered partitions $\tilde P$, we
-define
-\begin{align}
-  O_{\tilde S} \equiv \bigotimes_{O\in\tilde S} O,
-  &&
-  O_{\tilde P}
-  \equiv \bigotimes_{\tilde p\in\tilde P} \prod_{O\in\tilde p} O,
-  &&
-  \O_{\tilde S} \equiv \prod_{O\in\tilde S} \col{O},
-  &&
-  \O_{\tilde P}
-  \equiv \prod_{\tilde p\in\tilde P}
-  \col{\textstyle\prod_{O\in\tilde p} O},
-\end{align}
-where products over elements of $\tilde S,\tilde P,\tilde p$ are taken
-in order according to the elements of these ordered lists, e.g.~if
-$\tilde S=\p{O_1,O_2,\cdots,O_{\abs{S}}}$, then
-$\prod_{O\in\tilde S}\equiv O_1O_2\cdots O_{\abs{S}}$.  These
-definitions allow us to expand
-\begin{align}
-  \O_{\tilde S} \EQFS
-  \sum_{\substack{\t{ordered}\\\t{partitions}\\\tilde P~\t{of}~\tilde S}}
-  N!_{\abs{\tilde P}} O_{\tilde P},
-\end{align}
-in analogy to \eqref{eq:collective_to_multi_full}.
-
-While the expansions in \eqref{eq:multi_to_collective_full} and
-\eqref{eq:multi_to_collective_full_simp} closely resemble each other,
-in practice the latter involves less computational overhead due to the
-lack of symmetrization in the construction of $\O_{\tilde P}$.  There
-are, in principle, even more economical decompositions of a
-multi-local operators into collective opreators, for example
-\begin{align}
-  N!_2 X\otimes Y
-  &\EQFS \col{X}\,\col{Y} - \col{XY}, \\
-  N!_3 X\otimes Y \otimes Z
-  &\EQFS \col{X}\,\col{Y}\,\col{Z}
-  - \col{X}\,\col{YZ} - \col{Y}\,\col{XZ} - \col{Z}\,\col{XY}
-  + \col{ZXY} + \col{YXZ},
-\end{align}
-but these decompositions generally do not preserve the manifest
-symmetry of a multi-local operator $O_S$ under arbitrary permutations
-of $S$, and we do not know of any general prescription to arrive at
-such decompositions aside from solving systems of equations relating
-collective operators to multi-local operators.
-
-On a final note, we consider the special case that all operators
-$O\in S$ are identical, for which we provide the expansions
-\begin{align}
-  N!_2 X\otimes X \EQFS \col{X}^2 - \col{X^2},
-  &&
-  N!_3 X^{\otimes 3} \EQFS \col{X}^3
-  - 3 \col{X} \odot \col{X^2} + 2\col{X^3},
-\end{align}
-\begin{align}
-  N!_4 X^{\otimes 4}
-  \EQFS \col{X}^4 - 4 \col{X}^2 \odot \col{X^2}
-  - 2 \col{X} \, \col{X}^2 \, \col{X}
-  + 3 \p{\col{X^2}}^2 + 8 \col{X} \odot \col{X^3}
-  - 6 \col{X^4}.
-\end{align}
-If furthermore $X^2=1$, as with e.g.~the Pauli operators of SU(2),
-then
-\begin{align}
-  N!_2 X\otimes X \EQFS \col{X}^2 - N,
-  &&
-  N!_3 X^{\otimes 3} \EQFS \col{X}^3 - \p{3N-2} \col{X},
-\end{align}
-\begin{align}
-  N!_4 X^{\otimes 4} \EQFS \col{X}^4
-  - 2\p{3N-4} \col{X}^2 + 3 N \p{N-2}.
-\end{align}
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \section{Collective and multi-local operators in the fully symmetric
+%   manifold}
+% \label{sec:multi_to_collective}
+
+% Here we provide a decomposition multi-local operators into sums of
+% collective operators (and vice versa) within the fully symmetric
+% manifold of $N$ spins.  Multi-local operators are operators of the
+% form
+% \begin{align}
+%   O_S \equiv \bigotimes_{O\in S} O,
+%   \label{eq:multi_op_set}
+% \end{align}
+% where $S\equiv\set{O}$ is a set of single-spin operators $O$.  The
+% order of tensor factors in \eqref{eq:multi_op_set} does not matter
+% after projecting operators onto the fully symmetric manifold, as we
+% always will in this section.  While the decomposition of a collective
+% operator into a sum of multi-local operators is unique, the opposite
+% is not true: there is generally no unique decomposition of a
+% multi-local operator into a sum of collective operators.  We can,
+% however, enforce a canonical decomposition by choosing a set of
+% collective operators that obey the same symmetries as $O_S$, namely
+% invariance under arbitrary permutations of the single-spin operators
+% $O\in S$.
+
+% Even though $S$ contains $\abs{S}$ spins, a collective operator built
+% from the single-spin operators $O\in S$ may have multiple
+% (single-spin) operators acting on the same spin, resulting in a
+% $k$-local operator with $k<\abs{S}$.  We therefore {\it
+%   partition}\footnote{A partition $P\equiv\set{p}$ of $S$ is a set of
+%   mutually disjoint, non-empty subsets $p\subset S$ with
+%   $\bigcup_{p\in P}p=S$.} $S$ into subsets of operators that act on
+% the same spin.  For any partition $P\equiv\set{p}$ of $S$ (here
+% $p\subset S$ is a subset of operators $O\in S$ that address the same
+% spin), we then define the $\abs{P}$-local operator
+% \begin{align}
+%   O_P \equiv \bigotimes_{p\in P} \bigodot_{O\in p} O,
+% \end{align}
+% where $\bigodot_{x\in\X} x$ denotes a symmetrized product of all
+% elements in $\X$, i.e.
+% \begin{align}
+%   \bigodot_{x\in\X} x
+%   \equiv \f1{\abs{\X}!}
+%   \sum_{\substack{\t{permutations}\\\sigma~\t{of}~\X}}
+%   ~ \prod_{k=1}^{\abs{\X}} x_{\sigma_k},
+% \end{align}
+% with $\sigma_k$ the $k$-th element of the permutation $\sigma$ when
+% expressed in one-line notation; for example,
+% \begin{align}
+%   \bigodot_{j=1}^2 x_j &= x_1 \odot x_2 = \f12 \p{x_1x_2 + x_2x_1}, \\
+%   \bigodot_{j=1}^3 x_j
+%   &= \f1{3!} \p{x_1x_2x_3 + x_1x_3x_2 + x_2x_1x_3
+%     + x_2x_3x_1 + x_3x_1x_2 + x_3x_2x_1}.
+% \end{align}
+% The set of multi-local operators $\set{O_P:\t{partitions}~P~\t{of}~S}$
+% consists of all operators that
+% \begin{enumerate*}
+% \item can be built from ordinary products and tensor products of all
+%   single-spin operators $O\in S$, and
+% \item are invariant under arbitrary permutations of $S$.
+% \end{enumerate*}
+% For any set $S$ of single-spin operators and any partition $P$ of $S$,
+% we can similarly define the collective operators
+% \begin{align}
+%   \O_S \equiv \bigodot_{O\in S} \col{O},
+%   &&
+%   \O_P \equiv \bigodot_{p\in P} \col{\textstyle\bigodot_{O\in p} O},
+% \end{align}
+% where we underline a single-spin operator to denote its collective
+% version, i.e.~$\col{Q}\equiv\sum_j Q^{(j)}$, where $Q^{(j)}$ is an
+% operator that acts with the single-spin operator $Q$ on spin $j$ and
+% trivially on all other spins.  By construction, the set of collective
+% operators $\set{\O_P:\t{partitions}~P~\t{of}~S}$ is one-to-one to the
+% set of multi-local operators $\set{O_P:\t{partitions}~P~\t{of}~S}$,
+% and it is straightforward to work out that for any set $S$ of
+% single-spin operators,
+% \begin{align}
+%   \O_S
+%   \EQFS \sum_{\substack{\t{partitions}\\P~\t{of}~S}}
+%   N!_{\abs{P}} O_P,
+%   &&
+%   N!_k \equiv \prod_{j=0}^{k-1}\p{N-j}! = \f{N!}{\p{N-k}!},
+%   \label{eq:collective_to_multi_full}
+% \end{align}
+% where $\EQFS$ denotes equality under a restriction to the fully
+% symmetric manifold.  For any multi-local operator $O_S$ indexed by a
+% set $S$ of single-spin operators, we can collect expansions of the
+% form \eqref{eq:collective_to_multi_full} for all collective operators
+% built from non-empty subsets of $S$ into a system of linear equations
+% $\L$, and in turn solve $\L$ for $O_S$.  When
+% $\abs{S}\in\set{1,2,3,4}$ (and, by conjecture, for all $S$), by use of
+% a computer algebra system (namely, Mathematica) we find that
+% \begin{align}
+%   N!_{\abs{S}} O_S
+%   &\EQFS \sum_{\substack{\t{partitions}\\P~\t{of}~S}}
+%   \p{-1}^{\abs{S}-\abs{P}} \O_P \times \prod_{p\in P}\p{\abs{p}-1}!
+%   \label{eq:multi_to_collective_full} \\
+%   &\EQFS \f1{\abs{S}!} \sum_{\substack{\t{permutations}\\\Pi~\t{of}~S}}
+%   \sum_{\substack{\t{ordered}\\\t{partitions}\\\tilde P~\t{of}~\Pi}}
+%   \p{-1}^{\abs{S}-\abs{\tilde P}} \O_{\tilde P}
+%   \times \prod_{\tilde p\in\tilde P} \p{\abs{\tilde p}-1}!
+%   \label{eq:multi_to_collective_full_simp}
+% \end{align}
+% where we identify permutations $\Pi$ of $S$ with an ordered list
+% (tuple) $\p{O_1,O_2,\cdots,O_{\abs{S}}}$ of all single-spin operators
+% in $S$; and we enforce that partitions $\tilde P$ of an ordered list
+% $\Pi$, as well as parts $\tilde p\in\tilde P$, are lexicographically
+% ordered in accordance with the ordering of operators in $\Pi$.  If
+% $S=\set{X,Y,Z}$, for example, then the ordered partitions of the
+% trivial permutation $\Pi=\p{X,Y,Z}$ are
+% \begin{align}
+%   \p{\p{X,Y,Z}}, &&
+%   \p{\p{X,Y},\p{Z}}, &&
+%   \p{\p{X,Z},\p{Y}}, &&
+%   \p{\p{X},\p{Y,Z}}, &&
+%   \p{\p{X},\p{Y},\p{Z}}.
+% \end{align}
+% For ordered lists $\tilde S$ and ordered partitions $\tilde P$, we
+% define
+% \begin{align}
+%   O_{\tilde S} \equiv \bigotimes_{O\in\tilde S} O,
+%   &&
+%   O_{\tilde P}
+%   \equiv \bigotimes_{\tilde p\in\tilde P} \prod_{O\in\tilde p} O,
+%   &&
+%   \O_{\tilde S} \equiv \prod_{O\in\tilde S} \col{O},
+%   &&
+%   \O_{\tilde P}
+%   \equiv \prod_{\tilde p\in\tilde P}
+%   \col{\textstyle\prod_{O\in\tilde p} O},
+% \end{align}
+% where products over elements of $\tilde S,\tilde P,\tilde p$ are taken
+% in order according to the elements of these ordered lists, e.g.~if
+% $\tilde S=\p{O_1,O_2,\cdots,O_{\abs{S}}}$, then
+% $\prod_{O\in\tilde S}\equiv O_1O_2\cdots O_{\abs{S}}$.  These
+% definitions allow us to expand
+% \begin{align}
+%   \O_{\tilde S} \EQFS
+%   \sum_{\substack{\t{ordered}\\\t{partitions}\\\tilde P~\t{of}~\tilde S}}
+%   N!_{\abs{\tilde P}} O_{\tilde P},
+% \end{align}
+% in analogy to \eqref{eq:collective_to_multi_full}.
+
+% While the expansions in \eqref{eq:multi_to_collective_full} and
+% \eqref{eq:multi_to_collective_full_simp} closely resemble each other,
+% in practice the latter involves less computational overhead due to the
+% lack of symmetrization in the construction of $\O_{\tilde P}$.  There
+% are, in principle, even more economical decompositions of a
+% multi-local operators into collective opreators, for example
+% \begin{align}
+%   N!_2 X\otimes Y
+%   &\EQFS \col{X}\,\col{Y} - \col{XY}, \\
+%   N!_3 X\otimes Y \otimes Z
+%   &\EQFS \col{X}\,\col{Y}\,\col{Z}
+%   - \col{X}\,\col{YZ} - \col{Y}\,\col{XZ} - \col{Z}\,\col{XY}
+%   + \col{ZXY} + \col{YXZ},
+% \end{align}
+% but these decompositions generally do not preserve the manifest
+% symmetry of a multi-local operator $O_S$ under arbitrary permutations
+% of $S$, and we do not know of any general prescription to arrive at
+% such decompositions aside from solving systems of equations relating
+% collective operators to multi-local operators.
+
+% On a final note, we consider the special case that all operators
+% $O\in S$ are identical, for which we provide the expansions
+% \begin{align}
+%   N!_2 X\otimes X \EQFS \col{X}^2 - \col{X^2},
+%   &&
+%   N!_3 X^{\otimes 3} \EQFS \col{X}^3
+%   - 3 \col{X} \odot \col{X^2} + 2\col{X^3},
+% \end{align}
+% \begin{align}
+%   N!_4 X^{\otimes 4}
+%   \EQFS \col{X}^4 - 4 \col{X}^2 \odot \col{X^2}
+%   - 2 \col{X} \, \col{X}^2 \, \col{X}
+%   + 3 \p{\col{X^2}}^2 + 8 \col{X} \odot \col{X^3}
+%   - 6 \col{X^4}.
+% \end{align}
+% If furthermore $X^2=1$, as with e.g.~the Pauli operators of SU(2),
+% then
+% \begin{align}
+%   N!_2 X\otimes X \EQFS \col{X}^2 - N,
+%   &&
+%   N!_3 X^{\otimes 3} \EQFS \col{X}^3 - \p{3N-2} \col{X},
+% \end{align}
+% \begin{align}
+%   N!_4 X^{\otimes 4} \EQFS \col{X}^4
+%   - 2\p{3N-4} \col{X}^2 + 3 N \p{N-2}.
+% \end{align}
 

### 2020-01-31 collective_sun/sun_perturbation.tex
add diagrammatic calculation of single- and two-body operator products start calculation of three two-body Ising-like operators,   which will be necessary to compute matrix elements of a perturbation in the asymmetric manifolds

@@
 
-% for drawing set (Venn) diagrams
-\usepackage{tikz}
-\tikzset{
-  baseline = (current bounding box.center)
-}
-\usetikzlibrary{math}
+%%% figures
+\usepackage{graphicx} % for figures
+\graphicspath{{./figures/}} % set path for all figures
+\usepackage[export]{adjustbox} % for vertical alignment in math
+\newcommand{\diagram}[1]
+{\,\includegraphics[valign=c]{diagrams/#1.pdf}\,}
 
@@
 where $p,q$ index individual spins; $v_{Xp}$ and $w_{Opq}$ are real
-numbers; $X$ is a trace-zero single-body operator on $\H_n$; $O$ is a
-trace-zero two-body operator on $\H_n\otimes\H_n$ that obeys pair-wise
+numbers; $X$ is a trace-zero single-spin operator on $\H_n$; $O$ is a
+trace-zero two-spin operator on $\H_n\otimes\H_n$ that obeys pair-wise
 permutational symmetry, i.e.~$O_{pq}=O_{qp}$; $\v v_X$ and $\v X$ are
@@
 only on the coefficients $\v v_X^\Delta$, and is entirely independent
-of the state $\ket\psi$ or the choice of a trace-zero single-body
+of the state $\ket\psi$ or the choice of a trace-zero single-spin
 operators $X$ used to build $\V_1^\Delta$.  Finding operators
@@
 $\v v_X$ onto the $\Delta$-eigenspace of the matrix $\m h-\diag\v h$;
-and for any single-body operator $Z$, we define the collective operator
+and for any single-spin operator $Z$, we define the collective
+operator
 \begin{align}
@@
 \end{align}
-are an average of $w_{Opq}$ and a collective version of the two-body
-operator $O_{pq}$.  If the two-body operator $O=X\otimes Y$, then
+are an average of $w_{Opq}$ and a collective version of the two-spin
+operator $O_{pq}$.  If the two-spin operator $O=X\otimes Y$, then
 \begin{align}
@@
 \end{align}
-where $O$ is an $M$-body operator that is invariant under arbitrary
+where $O$ is an $M$-spin operator that is invariant under arbitrary
 permutation of its tensor factors; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is
@@
 where without loss of generality we consider a single-body
-perturbation $\V_1$ built out of only one single-body operator $X$.
+perturbation $\V_1$ built out of only one single-spin operator $X$.
 The sum in \eqref{eq:H_V1_psi_start} has terms with $k\in\set{p,q}$,
@@
 where without loss of generality we consider an $M$-body perturbation
-$\V_M$ built out of only one $M$-body operator $O$.  The sum in
+$\V_M$ built out of only one $M$-spin operator $O$.  The sum in
 \eqref{eq:H_VM_psi_start} has terms with $p,q\in k$, terms with
@@
 
+\subsection{The general case}
+
 Given a system of $N$ spins, we wish to project a product of
@@
 $k\in\C_N\p{M_j}$ is a choice of $M_j$ distinct spins; $O_j\p{k}$
-denotes the action of an $M_j$-body operator on spins $k$ that is
+denotes the action of an $M_j$-spin operator on spins $k$ that is
 invariant under permutations of $k$; and $w_j\p{k}$ is a scalar
@@
 tensors $w_j,O_j$ with $j\in S\subset\ZZ_p$.  For example, a term of
-the form $w_1\p{a,b,c,d} w_2\p{b,c,d} w_3\p{b}$ with distinct indices
-$\p{a,b,c,d}\in\C_N\p{4}$ would have
+the form $w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}$ with distinct
+indices $\p{a,b,c,d,e}\in\C_N\p{5}$ would have
 \begin{align}
@@
   &&
-  g_{\set{1,2}} = 2,
+  g_{\set{1,3}} = 1,
+  &&
+  g_{\set{2,3}} = 2,
 \end{align}
@@
 \begin{align}
-  \begin{tikzpicture}]
-    \tikzmath{
-      \scale = 0.7;
-      \rr = 2.5 * \scale;
-      \aa = \scale;
-      \bb = sqrt(3) * \aa;
-      \cc = 2 * \aa;
-      \topY = \cc + \rr;
-      \crossLeftRightUpper = sqrt(\rr*\rr - \bb*\bb) - \aa;
-      \dotTopY = ( \crossLeftRightUpper + \topY ) / 2;
-      \dotTopLeftHeight = \crossLeftRightUpper - \scale/5;
-      \dotTopLeftXFst = -\rr*0.48 - \scale/3;
-      \dotTopLeftXSnd = -\rr*0.48 + \scale/3;
-      \labelDist = \scale * 0.8;
-      \labelTopY = \topY + \labelDist;
-      \labelBottomX = \labelTopY * sqrt(3) / 2;
-      \labelBottomY = \labelTopY / 2;
-    }
-    \draw (0, \cc em) circle [radius = \rr em];
-    \draw (-\bb em, -\aa em) circle [radius = \rr em];
-    \draw (+\bb em, -\aa em) circle [radius = \rr em];
-    \node[draw,fill,shape=circle,scale=0.3] at (0,0) {};
-    \node[draw,fill,shape=circle,scale=0.3] at (0, \dotTopY em) {};
-    \node[draw,fill,shape=circle,scale=0.3]
-    at (\dotTopLeftXFst em, \dotTopLeftHeight em) {};
-    \node[draw,fill,shape=circle,scale=0.3]
-    at (\dotTopLeftXSnd em, \dotTopLeftHeight em) {};
-    \node at (0, \labelTopY em) {1};
-    \node at (-\labelBottomX em, -\labelBottomY em) {2};
-    \node at (+\labelBottomX em, -\labelBottomY em) {3};
-  \end{tikzpicture},
+  \diagram{example},
   \label{eq:venn_diagram}
@@
 where $g_S$ is determined the number of dots at the intersection of
-the circles $j\in S$.  We denote an assignment of distinct indices
+circles $j\in S$.  We denote an assignment of distinct indices
 according to choices of $g_S$ for all $S\in\PP\p{\ZZ_p}$ by $g$, where
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Product of two single-body operators}
+\subsection{Two single-body operators}
 \label{sec:PXYP}
@@
 \eqref{eq:perturbations} requires us to simplify a product of the form
-\begin{align}
-  \P_0 \sp{\sum_{p,q} v_{Xp} v_{Yq} X_p Y_q} \P_0
-  \EQFS \sum_p v_{Xp} v_{Yp} X_1 Y_1
-  + \sum_{p\ne q} v_{Xp} v_{Yq} X_1 Y_2,
+
+\begin{align}
+  \P_0 \sp{\sum_{p,q} v_p w_q X_p Y_q} \P_0
+  \EQFS \diagram{single_body_0} X_1 Y_2
+  + \diagram{single_body_1} X_1 Y_1,
   \label{eq:PXYP_start}
 \end{align}
-where $X,Y$ are single-body operators; and the coefficients
-$v_{Xp},v_{Yq}$ satisfy \eqref{eq:cond_1} with eigenvalue
-$\Delta\ne0$, which implies that they are mean-zero.  The first sum on
-the right of \eqref{eq:PXYP_start} is then
-\begin{align}
-  \sum_p v_{Xp} v_{Yp} X_1 Y_1
-  \EQFS \f1N \v v_X \c \v v_Y\, \col{XY},
-  &&
-  \col{Z} \equiv \sum_p Z_p.
-\end{align}
-To simplify the second sum on the right of \eqref{eq:PXYP_start}, we
-use the fact that $\v v$ is mean-zero to write
-\begin{align}
-  \sum_{p\ne q} v_{Xp} v_{Yq}
-  = \sum_{p,q} v_{Xp} v_{Yq} - \sum_{p=q} v_{Xp} v_{Yq}
-  = - \v v_X \c\v v_Y,
-\end{align}
-and we expand
+where $p,q$ index individual spins; $X,Y$ are single-spin operators;
+and the coefficients $v_p,w_q$ satisfy \eqref{eq:cond_1} with
+eigenvalue $\Delta\ne0$, which implies that they are mean-zero.
+Writing out
+\begin{align}
+  \diagram{single_body_1}
+  \equiv \sum_p v_p w_p
+  = \v v\c\v w,
+\end{align}
+\begin{align}
+  \diagram{single_body_0}
+  \equiv \sum_{p\ne q} v_p w_q
+  = \sum_{p,q} v_p w_q - \sum_{p=q} v_p w_q
+  = - \v v \c\v w,
+\end{align}
+we thus have
+\begin{align}
+  \sum_{p,q} v_p w_q X_p Y_q
+  \EQFS - \v v\c\v w \p{X_1 Y_2 - X_1 Y_1},
+\end{align}
+This result is also simple to write in terms of collective operators,
+as
+\begin{align}
+  \col{X Y} \EQFS N X_1 Y_1,
+  &&
+  \col{Z} \equiv \sum_p Z_p,
+\end{align}
 \begin{align}
@@
 \end{align}
-which implies that
-\begin{align}
-  \sum_{p\ne q} v_{Xp} v_{Yq} X_1 Y_2
-  \EQFS - \v v_X\c\v v_Y \p{\col{X}\,\col{Y} - \col{XY}}.
-\end{align}
-Altogether, we have that
-\begin{align}
-  \sum_{p,q} v_{Xp} v_{Yq} X_p Y_q
-  \EQFS - \f{\v v_X \c\v v_Y}{N\p{N-1}}
+so
+\begin{align}
+  \sum_{p,q} v_p w_q X_p Y_q
+  \EQFS - \f{\v v \c\v w}{N\p{N-1}}
   \p{\col{X}\,\col{Y} - N \col{XY}}.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Product of two two-body operators}
+\subsection{Two two-body operators}
 \label{sec:POQP}
@@
 \begin{multline}
-  \P_0 \sp{\sum_{\substack{k<l\\p<q}}
-    w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq}} \P_0
-  \EQFS \sum_{p<q} w_{Opq} w_{Qpq} O_{1,2} Q_{1,2}
-  + \sum_{\substack{k,p,q\\k\ne p\ne q\ne k}}
-  w_{Okp} w_{Qkq} O_{1,2} Q_{1,3} \\
-  + \sum_{\substack{k<\ell\\p<q\\\set{k,\ell}\ne\set{p,q}}}
-  w_{Ok\ell} w_{Qpq} O_{1,2} Q_{3,4},
+  \P_0 \sp{\sum_{k<\ell} \sum_{p<q}
+    v_{k\ell} w_{pq} O_{k\ell} Q_{pq}} \P_0 \\
+  \EQFS \diagram{two_body_0} O_{1,2} Q_{3,4}
+  + \diagram{two_body_1} O_{1,2} Q_{1,3}
+  + \diagram{two_body_2} O_{1,2} Q_{1,2},
   \label{eq:POQP_start}
 \end{multline}
-where $O$ and $Q$ are permutationally-symmetric two-body operators;
-and the coefficients $w_{Ok\ell},w_{Qpq}$ satisfy \eqref{eq:cond_2}
-with eigenvalue $\Delta\ne0$, which implies that they are mean-zero.
-Defining
+where $k,\ell,p,q$ index individual spins; $O$ and $Q$ are
+permutationally-symmetric two-spin operators; and the coefficients
+$v_{k\ell},w_{pq}$ satisfy \eqref{eq:cond_2} with eigenvalue
+$\Delta\ne0$, which implies that they are mean-zero.  Defining
 \begin{align}
@@
 \end{align}
-we can write the first sum on the right of \eqref{eq:POQP_start} as
-\begin{align}
-  \sum_{p<q} w_{Opq} w_{Qpq} = \v{\m w}_O \c\v{\m w}_Q,
-\end{align}
-and, using the fact that $w_{Opp}=0$, we can write the second sum on
-the right of \eqref{eq:POQP_start} as
-\begin{align}
-  \sum_{\substack{k,p,q\\k\ne p\ne q\ne k}} w_{Okp} w_{Qkq}
-  = \sum_{k,p} w_{Okp} \sum_{q\ne p} w_{Okq}
-  = \sum_{k,p} w_{Okp} \p{w_{Qk} - w_{Qkp}}
-  = \v w_O \c \v w_Q - 2 \v{\m w}_O \c \v{\m w}_Q.
-\end{align}
-Using additionally the fact that $w_{Opq},w_{Qpq}$ are mean-zero, we
-can similarly simplify
-\begin{align}
-  \sum_{\substack{p<q\\k<\ell\\\set{k,\ell}\ne\set{p,q}}}
-  w_{Ok\ell} w_{Qpq} = -\v w_O \c \v w_Q + \v{\m w}_O \c \v{\m w}_Q.
+we can write out explicitly
+\begin{align}
+  \diagram{two_body_2}
+  \equiv \sum_{p<q} v_{pq} w_{pq} = \v{\m v} \c\v{\m w},
+\end{align}
+and, using the fact that $v_{pp}=w_{pp}=0$,
+\begin{align}
+  \diagram{two_body_1}
+  \equiv \sum_{\substack{k,p,q\\k\ne p\ne q\ne k}} v_{kp} w_{kq}
+  = \sum_{k,p} v_{kp} \sum_{q\ne p} w_{kq}
+  = \sum_{k,p} v_{kp} \p{w_{k} - w_{kp}}
+  = \v v \c \v w - 2 \v{\m v} \c \v{\m w}.
+\end{align}
+Using additionally the fact that $v_{pq},w_{pq}$ are mean-zero, with
+some more work we can work out that
+\begin{align}
+  \diagram{two_body_0}
+  \equiv \sum_{k<\ell} \sum_{\substack{p<q\\p,q\notin\set{k,\ell}}}
+  v_{k\ell} w_{pq} = -\v v \c \v w + \v{\m v} \c \v{\m w}.
 \end{align}
@@
 \begin{multline}
-  \sum_{\substack{k<l\\p<q}} w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq}
-  \EQFS \v{\m w}_O \c \v{\m w}_Q O_{1,2} Q_{1,2}
-  + \p{\v w_O\c\v w_Q - 2 \v{\m w}_O \c \v{\m w}_Q} O_{1,2} Q_{1,3} \\
-  - \p{\v w_O \c \v w_Q - \v{\m w}_O \c \v{\m w}_Q} O_{1,2} Q_{3,4}.
+  \sum_{k<\ell} \sum_{p<q} v_{k\ell} w_{pq} O_{k\ell} Q_{pq} \\
+  \EQFS - \p{\v v \c \v w - \v{\m v} \c \v{\m w}} O_{1,2} Q_{3,4}
+  + \p{\v v\c\v w - 2 \v{\m v} \c \v{\m w}} O_{1,2} Q_{1,3}
+  + \v{\m v} \c \v{\m w}\, O_{1,2} Q_{1,2}.
   \label{eq:POQP}
 \end{multline}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Three two-body Ising-like operators}
+
+Here we simplify a product of three two-body operators of a particular
+form projected onto the fully symmetric manifold:
+\begin{align}
+  \X \equiv \P_0 \sp{\sum_{k<\ell} \sum_{p<q} \sum_{r<s}
+    u_{k\ell} v_{pq} w_{rs} X_k X_\ell X_p X_q X_r X_s} \P_0,
+  &&
+  X^2 = 1,
+\end{align}
+where $k,\ell,p,q,r,s$ index individual spins; $X$ is a single-spin
+operator; and $u_{k\ell},v_{pq},w_{rs}$ are arbitrary scalar
+coefficients.  Collecting terms according to the number of $X$
+operators that remain after accounting for $X^2=1$, we find that
+\begin{align}
+  \X \EQFS A_6 X^{\otimes 6} + A_4 X^{\otimes 4} + A_2 X^{\otimes 2},
+\end{align}
+with
+\begin{align}
+  A_6 \equiv \diagram{triple_6},
+  &&
+  A_4 \equiv \diagram{triple_4_1} + \diagram{triple_4_0},
+  \label{eq:triple_64}
+\end{align}
+\begin{align}
+  A_2 \equiv \diagram{triple_2_0} + \diagram{triple_2_1}
+  + \diagram{triple_2_2} + \diagram{triple_2_3},
+  \label{eq:triple_2}
+\end{align}
+where a diagram with unlabeled circles implicitly contains a sum over
+all label assignments that result in a distinct diagram, e.g.
+\begin{align}
+  \diagram{triple_6} \equiv \diagram{triple_6_uvw},
+  \label{eq:trible_uvw}
+\end{align}
+\begin{align}
+  \diagram{triple_2_3} \equiv \diagram{triple_2_3_uvw}
+  + \diagram{triple_2_3_vwu} + \diagram{triple_2_3_wuv}.
+  \label{eq:trible_uvw_3}
+\end{align}
+The first diagram in each of $A_6,A_4,A_2$ as written in
+\eqref{eq:triple_64}, \eqref{eq:triple_2} thus has only one assignment
+of labels, as in \eqref{eq:trible_uvw}, while all other unlabeled
+diagrams have three assignments, as in \eqref{eq:trible_uvw_3}.
+
+\vspace{5cm}
 

### 2020-02-01 collective_sun/sun_perturbation.tex
expressed triple product in terms of collective operators

@@
 \end{align}
-we can write out explicitly
+for $m\in\set{v,w}$, we can write out explicitly
 \begin{align}
@@
 operator; and $u_{k\ell},v_{pq},w_{rs}$ are arbitrary scalar
-coefficients.  Collecting terms according to the number of $X$
+coefficients with $m_{pq}=m_{qp}$ and $m_{pp}=0$ for each of
+$m\in\set{u,v,w}$.  Collecting terms according to the number of $X$
 operators that remain after accounting for $X^2=1$, we find that
@@
   \X \EQFS A_6 X^{\otimes 6} + A_4 X^{\otimes 4} + A_2 X^{\otimes 2},
+  \label{eq:triple_multi}
 \end{align}
@@
   \diagram{triple_6} \equiv \diagram{triple_6_uvw},
-  \label{eq:trible_uvw}
+  \label{eq:triple_uvw}
 \end{align}
@@
   + \diagram{triple_2_3_vwu} + \diagram{triple_2_3_wuv}.
-  \label{eq:trible_uvw_3}
+  \label{eq:triple_uvw_3}
 \end{align}
@@
 \eqref{eq:triple_64}, \eqref{eq:triple_2} thus has only one assignment
-of labels, as in \eqref{eq:trible_uvw}, while all other unlabeled
-diagrams have three assignments, as in \eqref{eq:trible_uvw_3}.
+of labels, as in \eqref{eq:triple_uvw}, while all other unlabeled
+diagrams have three assignments, as in \eqref{eq:triple_uvw_3}.  We
+can write \eqref{eq:triple_multi} in terms of collective operators as
+\begin{align}
+  \X \EQFS
+  \tilde A_6 \col{X}^6 + \tilde A_4 \col{X}^4 + \tilde A_2 \col{X}^2,
+  \label{eq:triple_col}
+\end{align}
+where we have neglected scalar terms without physical consequence, and
+\begin{align}
+  \tilde A_6 \equiv \f{A_6}{N!_6},
+  &&
+  \tilde A_4 \equiv \f{A_4}{N!_4} - 5\p{3N-8} \f{A_6}{N!_6},
+\end{align}
+\begin{align}
+  \tilde A_2 \equiv \f{A_2}{N!_2} - 2\p{3N-4} \f{A_4}{N!_4}
+  + \p{45N^2-210N+184} \f{A_6}{N!_6}.
+\end{align}
 

### 2020-02-02 collective_sun/scripts/check_multibody_analytics.py
added code to check multibody eigenvalue problem

@@ -104,7 +104,7 @@ def act(op, indices):
     return np.reshape(op, (dim**spins, dim**spins))
 
 ##########################################################################################
-# construct projector onto the fully symmetric manifold
+# methods to construct the product of multi-body operators from set diagrams
 ##########################################################################################
 
 # build random multi-body operators, each defined by a tensor and a "base" operator
@@ -175,5 +175,43 @@ for set_diagram in set_diagrams(ranks):
 
     simp_op += diagram_coefficient * ( sym_proj @ diagram_op @ sym_proj )
 
-# check whether the exact and simplified products of multi-body operators are equal
+# verify that the exact and simplified products of multi-body operators are equal
 print(abs(simp_op-exact_op).max() < cutoff)
+
+##########################################################################################
+# methods to verify the multi-body eigenvalue problem
+##########################################################################################
+
+# build random SU(n)-symmetric interaction Hamiltonian
+sun_coeffs = random_tensor(2)
+swap = sum( ( lambda op : np.kron(op.transpose(),op) )
+            ( np.kron(unit_vector(dim,nu),
+                      unit_vector(dim,mu)) )
+            for mu in range(dim) for nu in range(dim) )
+sun_interactions = sum( sun_coeffs[pp,qq] * act(swap,[pp,qq])
+                        for pp in range(spins) for qq in range(pp) )
+
+# build objects that appear in the multi-body eigenvalue problem
+sym_energy = sun_coeffs.sum()/2
+sun_coeff_vec = sum(sun_coeffs)
+def _coeff_mod(spin_choice):
+    pos = sum( sun_coeffs[pp,qq] for pp,qq in it.combinations(spin_choice, 2) )
+    neg = sum( sun_coeff_vec[pp] for pp in spin_choice )
+    return 2 * pos - neg
+def _coeff_zip_tensor(tensor):
+    return sum( np.tensordot(sun_coeffs, tensor, axes = [ 1, jj ])
+                for jj in range(tensor.ndim) )
+
+# verify the multi-body eigenvalue problem for every multi-body operator that we built
+for tensor, base_op, full_op in zip(tensors, base_ops, full_ops):
+    exact = sun_interactions @ full_op @ sym_proj
+
+    _coeff_zip = _coeff_zip_tensor(tensor)
+    remaining_op = sum( ( _coeff_mod(idx) * tensor[idx] + _coeff_zip[idx] )
[…]

### 2020-02-02 collective_sun/sun_perturbation.tex
added code to check multibody eigenvalue problem
rearranged some math

@@
   = E_0 \V_2 \ket\psi
-  + \sum_O \sum_{p<q} \sp{\bar h_{pq} w_{Opq}
-    + \sum_k \p{h_{kp} w_{Okq} w_{Okq} + h_{kq} w_{Opk}}}
+  + \sum_O \sum_{p<q}
+  \sp{\sum_k \p{h_{kp} w_{Okq} w_{Okq} + h_{kq} w_{Opk}}
+  - \bar h_{pq} w_{Opq}}
   O_{pq} \ket\psi,
@@
 \begin{align}
-  \bar h_{pq} \equiv 2 h_{pq} - h_p - h_q,
+  \bar h_{pq} \equiv h_p + h_q - 2 h_{pq},
   &&
@@
 \begin{align}
-  \check{\m h} \equiv \sum_{p<q} \bar h_{pq} \op{\set{p,q}}
-  + \sum_{p<q} \ket{\set{p,q}} \sum_{k\notin\set{p,q}}
-  \p{h_{kp} \bra{\set{k,q}} + h_{kq} \bra{\set{p,k}}}.
+  \check{\m h}
+  \equiv \sum_{p<q} \ket{\set{p,q}} \sp{\sum_{k\notin\set{p,q}}
+    \p{h_{kp} \bra{\set{k,q}} + h_{kq} \bra{\set{p,k}}}
+    - \bar h_{pq} \bra{\set{p,q}}}.
   \label{eq:h_super_mat}
@@
   X_p \ket\psi.
+  \label{eq:H_V1_psi_end}
 \end{align}
@@
   \sum_{\substack{\p{p,q}\in\C_N\p{2}\\p,q\in k~\t{or}~p,q\notin k}} h_{pq}
-  = E_0 + \bar h_k,
-  &&
-  \bar h_k \equiv 2\sum_{\p{p,q}\subset k} h_{pq} - \sum_{p\in k} h_p,
-  &&
-  h_p \equiv \sum_q h_{pq},
-\end{align}
-where the sum over $\p{p,q}\subset k$ denotes a sum over distinct
-choices of two elements in $k$.  The terms in
-\eqref{eq:H_VM_psi_start} with only one of $p\in k$ or $q\in k$,
-meanwhile, are
+  = E_0 - \bar h_k,
+  &&
+  \bar h_k \equiv \sum_{p\in k}\p{h_p - \sum_{q\in k}h_{pq}},
+  &&
+  h_p \equiv \sum_q h_{pq}.
+\end{align}
+The terms in \eqref{eq:H_VM_psi_start} with only one of $p\in k$ or
+$q\in k$, meanwhile, are
 \begin{align}
@@
   = \sum_{\substack{p\in\ZZ_N\\p\notin k}}
-  \p{h_{pk_1} w_{pk_2k_3\cdots k_M}
-    + h_{pk_2} w_{k_1pk_3\cdots k_M}
-    + \cdots + h_{pk_M} w_{k_1k_2\cdots k_{M-1} p}},
+  \p{h_{k_1 p} w_{pk_2k_3\cdots k_M}
+    + h_{k_2 p} w_{k_1pk_3\cdots k_M}
+    + \cdots + h_{k_M p} w_{k_1k_2\cdots k_{M-1} p}},
 \end{align}
@@
   = E_0 \V_M \ket\psi + \sum_{k\in\C_N\p{M}}
-  \sp{\bar h_k w_k + \p{h\circ w}_k} O_k \ket\psi.
-\end{align}
-If $M=1$, then $\bar h_k=-h_k$ and
-$\p{h\circ w}_k=\sum_\ell h_{k\ell}w_\ell$, so
+  \sp{\p{h\circ w}_k - \bar h_k w_k} O_k \ket\psi.
+\end{align}
+If $M=1$, then $\p{h\circ w}_k=\sum_\ell h_{k\ell}w_\ell$ and
+$\bar h_k=h_k$, so
 \begin{align}
@@
   = E_0 \V_1 \ket\psi + \sum_{k\in\ZZ_N}
-  \p{\sum_\ell h_{k\ell} w_\ell - h_k w_k} O_k \ket\psi.
-\end{align}
-If $M=2$, then
-$\p{h\circ w}_{pq}=\sum_k\p{h_{kp} w_{kq}+h_{kq}w_{pk}}$, so
+  \p{\sum_{\ell\in\ZZ_N} h_{k\ell} w_\ell - h_k w_k} O_k \ket\psi,
+\end{align}
+which is precisely what we found in \eqref{eq:H_V1_psi_end}.  If
+$M=2$, then $\p{h\circ w}_{pq}=\sum_k\p{h_{kp} w_{kq}+h_{kq}w_{pk}}$,
+so
 \begin{align}
   H_0 \V_2 \ket\psi
-  = E_0 \V_2 \ket\psi + \sum_{p<q}
-  \sp{\bar h_{pq} w_{pq} + \sum_k\p{h_{kp} w_{kq} + h_{kq} w_{pk}}}
+  = E_0 \V_2 \ket\psi + \sum_{\p{p,q}\in\C_N\p{2}}
+  \sp{\sum_{k\in\ZZ_N}\p{h_{kp} w_{kq} + h_{kq} w_{pk}}
+    - \bar h_{pq} w_{pq}}
   O_k \ket\psi,
 \end{align}
-where $\bar h_{pq}=2h_{pq}-h_p-h_q$ with $h_p\equiv\sum_qh_{pq}$.
+where $\bar h_{pq}=h_p+h_q-2h_{pq}$ with $h_p\equiv\sum_qh_{pq}$.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Analytical results in translationally invariant systems}
+\section{Making use of translational invariance}
 \label{sec:trans_inv}
@@
 \begin{align}
-  \bar h_{pq} w_{pq}
-  + \sum_k \p{h_{pk} w_{kq} + w_{pk} h_{kq}}
+  \sum_k \p{h_{pk} w_{kq} + w_{pk} h_{kq}}
+  - \bar h_{pq} w_{pq}
   = \Delta w_{pq},
   &&
-  \bar h_{pq} \equiv 2h_{pq} - h_p - h_q,
+  \bar h_{pq} \equiv h_p + h_q - 2h_{pq},
   &&
@@
 \begin{align}
-  \bar{\m h} * \m w + \P_{\t{off-diag}}\p{\m h\c\m w + \m w\c\m h}
+  \P_{\t{off-diag}}\p{\m h\c\m w + \m w\c\m h} - \bar{\m h} * \m w
   = \Delta \m w,
@@
 where $\m h$, $\bar{\m h}$, and $\m w$ are matrices of all $h_{pq}$,
-$\bar h_{pq}$ and $w_{pq}$; $\bar{\m h}*\m w$ denotes an element-wise
-(Kronecker) product of $\bar{\m h}$ and $\m w$, i.e.~with matrix
-elements $\bar h_{pq}w_{pq}$; and
+$\bar h_{pq}$ and $w_{pq}$;
 $\P_{\t{off-diag}}\p{\m m}\equiv \m m-\sum_p\op{p}\m m\op{p}$ denotes
 the projection of $\m m$ onto the space of strictly off-diagonal
-matrices.  In order to find solutions to \eqref{eq:cond_2_ref} when
-the interaction Hamiltonian $H_0$ is translationally invariant, in
-which case $h_{pq}$ and $\bar h_{pq}$ depend only on the separation
-$\pm\p{p-q}$, we expand all matrices in a basis that the respects
-translational symmetry.  To this end, for all lattice displacements
-$d\in\ZZ_L^D$ and wavenumbers $k\in\ZZ_L^D\times2\pi/L$ we define the
-{\it cycle matrices}
+matrices; and $\bar{\m h}*\m w$ denotes an element-wise (Kronecker)
+product of $\bar{\m h}$ and $\m w$, i.e.~with matrix elements
+$\bar h_{pq}w_{pq}$.  In order to find solutions to
+\eqref{eq:cond_2_ref} when the interaction Hamiltonian $H_0$ is
+translationally invariant, in which case $h_{pq}$ and $\bar h_{pq}$
+depend only on the separation $\pm\p{p-q}$, we expand all matrices in
+a basis that the respects translational symmetry.  To this end, for
+all lattice displacements $d\in\ZZ_L^D$ and wavenumbers
+$k\in\ZZ_L^D\times2\pi/L$ we define the {\it cycle matrices}
 \begin{align}
@@
 \begin{align}
-  \sum_{d\ne0} \bar h_{d,0} \tilde w_{dk} \m m_{dk}
-  + \sum_{\substack{c,d\\c+d\ne0}}
+  \sum_{\substack{c,d\\c+d\ne0}}
   h_{c,0} \tilde w_{dk} \p{1 + e^{-ic\c k}} \m m_{c+d,k}
+  - \sum_{d\ne0} \bar h_{d,0} \tilde w_{dk} \m m_{dk}
   = \Delta \sum_{d\ne0} \tilde w_{dk} \m m_{dk},
@@
 \begin{align}
-  \sum_{d\ne0} \bar h_{d,0} \tilde w_{dk} \m m_{dk}
-  + \sum_{c,d\ne0} h_{d-c,0} \tilde w_{ck}
+  \sum_{c,d\ne0} h_{d-c,0} \tilde w_{ck}
   \p{1 + e^{-i\p{d-c}\c k}} \m m_{dk}
+  - \sum_{d\ne0} \bar h_{d,0} \tilde w_{dk} \m m_{dk}
   = \Delta \sum_{d\ne0} \tilde w_{dk} \m m_{dk},
@@
   \ut{\check{\m h}}_k
-  \equiv \sum_{d\ne0} \bar h_{d,0} \op{d}
-  + \sum_{c,d\ne0} h_{d-c,0} \p{1+e^{-i\p{d-c}\c k}} \op{d}{c},
+  \equiv \sum_{c,d\ne0} h_{d-c,0} \p{1+e^{-i\p{d-c}\c k}} \op{d}{c}
+  - \sum_{d\ne0} \bar h_{d,0} \op{d},
   &&
@@
   \ut{\check{\m h}}_k
-  &= \sum_{d\ne0} \sp{-2+2\p{N-1}} \op{d}
-  - \sum_{\substack{c,d\ne0\\c\ne d}} \p{1+e^{-i\p{d-c}\c k}} \op{d}{c} \\
-  &= 2\p{N-1} \1_{N-1}
-  - \sum_{c,d\ne0} \p{1+e^{-i\p{d-c}\c k}} \op{d}{c},
+  &= -\sum_{\substack{c,d\ne0\\c\ne d}}
+  \p{1+e^{-i\p{d-c}\c k}} \op{d}{c}
+  + \sum_{d\ne0} \sp{2\p{N-1}-2} \op{d} \\
+  &= -\sum_{c,d\ne0} \p{1+e^{-i\p{d-c}\c k}} \op{d}{c}
+  + 2\p{N-1} \1_{N-1},
 \end{align}
@@
   \ut{\check{\m h}}_k
-  = 2\p{N-1} \1_{N-1} -\p{N-1} \p{\op{\ut{0}} + \op{\ut{k}}}.
+  = 2 \p{N-1} \1_{N-1} - \p{N-1} \p{\op{\ut{0}} + \op{\ut{k}}}.
 \end{align}
@@
   \ut{\check{\m h}}_k
-  = 2\p{N-1} \1_{N-1} - \sum_{s\in\set{\pm}} \varepsilon_s \op{\ut{k}^s}
+  = 2 \p{N-1} \1_{N-1}
+  - \sum_{s\in\set{\pm}} \varepsilon_s \op{\ut{k}^s}
 \end{align}

### 2020-02-03 collective_sun/sun_perturbation.tex
rearranged appendices renamed the 'fully symmetric' manifold to the 'permutationally symmetric' manifold

@@
 all negative, then the ground-state manifold $\M_0$ of the interaction
-Hamiltonian $H_0$ consists of fully symmetric states that are
-simultaneous $+1$ eigenstates of all permutation operators $\Pi_{pq}$.
-We will adopt the restriction that all $h_{pq}<0$ throughout these
-notes, but note that this restriction can be relaxed to the assumption
-that the initial state of the spins is fully symmetric (e.g.~a
-spin-polarized state).  We will also assume throughout these notes
-that the fully symmetric manifold $\M_0$ is gapped away from all other
-states by an interaction energy difference that is large compared to
-any coupling between $\M_0$ and its orthogonal complement.  Power-law
-couplings of the form $h_{pq}=-h/\abs{p-q}^\alpha$ on a
-$D$-dimensional lattice, for example, yield a non-vanishing spectral
-gap in the thermodynamic limit when $\alpha\le D$ (see Appendix
-\ref{sec:gap}).
+Hamiltonian $H_0$ consists of permutationally symmetric states that
+are simultaneous $+1$ eigenstates of all permutation operators
+$\Pi_{pq}$.  We will adopt the restriction that all $h_{pq}<0$
+throughout these notes, but note that this restriction can be relaxed
+to the assumption that the initial state of the spins is
+permutationally symmetric (e.g.~a spin-polarized state).  We will also
+assume throughout these notes that the permutationally symmetric
+manifold $\M_0$ is gapped away from all other states by an interaction
+energy difference that is large compared to any coupling between
+$\M_0$ and its orthogonal complement.  Power-law couplings of the form
+$h_{pq}=-h/\abs{p-q}^\alpha$ on a $D$-dimensional lattice, for
+example, yield a non-vanishing spectral gap in the thermodynamic limit
+when $\alpha\le D$ (see Appendix \ref{sec:gap}).
 
@@
 interaction Hamiltonian $H_0$, with interaction energy $\Delta$ above
-that of fully symmetric manifold $\M_0$; that is, the interaction
-energy of states in the image of $\P_\Delta$ is $E_0+\Delta$, where
+that of permutationally symmetric manifold $\M_0$; that is, the
+interaction energy of states in the image of $\P_\Delta$ is
+$E_0+\Delta$, where
 \begin{align}
@@
 \end{align}
-is the interaction energy of fully symmetric states $\ket\psi\in\M_0$.
+is the interaction energy of permutationally symmetric states
+$\ket\psi\in\M_0$.
 
@@
 of $X$.  In order to compute the second order effective Hamiltonian
-$H_1^{(2)}$, we first choose an arbitrary fully symmetric state
-$\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_V1_psi}):
+$H_1^{(2)}$, we first choose an arbitrary permutationally symmetric
+state $\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_V1_psi}):
 \begin{align}
@@
 $\V_1^\Delta$ that generate eigenvectors of the interaction
-Hamiltonian $H_0$ when they act on fully symmetric states $\ket\psi$
-thus reduces to finding eigenvectors of the interaction matrix
-$\m h-\diag\v h$.
-
-If the fully symmetric manifold $\M_0$ is gapped away from all other
-states, then then a vector $\V_1^\Delta\ket\psi$ with $\Delta=0$ must
-lie within the fully symmetric manifold $\M_0$, which implies that the
-operator $\V_1^\Delta$ preserves the permutational symmetry of $\M_0$.
-Indeed, a constant vector $\v\1\equiv\p{1,1,1,\cdots}$ of ones
-satisfies the condition in \eqref{eq:cond_1} with eigenvalue $0$,
-which implies that all vectors $\v v$ satisfying \eqref{eq:cond_1}
-with $\Delta\ne0$ must be orthogonal to $\v\1$, and therefore
-mean-zero.
+Hamiltonian $H_0$ when they act on permutationally symmetric states
+$\ket\psi\in\M_0$ thus reduces to finding eigenvectors of the
+interaction matrix $\m h-\diag\v h$.
+
+If the permutationally symmetric manifold $\M_0$ is gapped away from
+all other states, then then a vector $\V_1^\Delta\ket\psi$ with
+$\Delta=0$ must lie within the permutationally symmetric manifold
+$\M_0$, which implies that the operator $\V_1^\Delta$ preserves the
+permutational symmetry of $\M_0$.  Indeed, a constant vector
+$\v\1\equiv\p{1,1,1,\cdots}$ of ones satisfies the condition in
+\eqref{eq:cond_1} with eigenvalue $0$, which implies that all vectors
+$\v v$ satisfying \eqref{eq:cond_1} with $\Delta\ne0$ must be
+orthogonal to $\v\1$, and therefore mean-zero.
 
@@
 where each operator $\V_1^\Delta$ generates a state with definite
-interaction energy $\Delta$ above that of the fully symmetric manifold
-$\M_0$.  We can therefore simplify
+interaction energy $\Delta$ above that of the permutationally
+symmetric manifold $\M_0$.  We can therefore simplify
 \begin{align}
@@
 In order to compute the second-order effective Hamiltonian
-$H_2^{(2)}$, as before we pick an arbitrary fully symmetric state
-$\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_VM_psi})
+$H_2^{(2)}$, as before we pick an arbitrary permutationally symmetric
+state $\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_VM_psi})
 \begin{align}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Diagnosing perturbed states with the interaction Hamiltonian}
+\section{Constructing eigenstates of multi-body perturbations}
 \label{sec:H_VM_psi}
@@
 In this section, we simplify the products of the form
-$H_0\V_M\ket\psi$ to arrive at the expansions in \eqref{eq:H_V1_psi}
-and \eqref{eq:H_V2_psi}.  Here $H_0$ is the SU($n$)-symmetric
-interaction Hamiltonian in \eqref{eq:ints}, $\ket\psi$ is a fully
-symmetric state, and $\V_M$ is an $M$-body perturbation of the form
+$H_0\V_M\ket\psi$ to arrive at expansions such as \eqref{eq:H_V1_psi}
+and \eqref{eq:H_V2_psi}, which define conditions under which a
+multi-body perturbation $\V_M$ generates eigenstates of the
+SU($n$)-symmetric interaction Hamiltonian $H_0$ in \eqref{eq:ints}
+when acting on a permutationally symmetric state $\ket\psi\in\M_0$.
+Here $\V_M$ is an $M$-body perturbation of the form
 \begin{align}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Operator products restricted to the fully symmetric manifold}
-\label{sec:sym_prod}
-
-\subsection{The general case}
-
-Given a system of $N$ spins, we wish to project a product of
-multi-body operators onto the fully symmetric manifold.  For $p$
-multi-body operators, the projection of such a product takes the form
-\begin{align}
-  \P_0 \sp{\prod_{j\in\ZZ_p}
-    \sum_{k\in\C_N\p{M_j}} w_j\p{k} O_j\p{k}} \P_0,
-  \label{eq:sym_prod_proj}
-\end{align}
-where $\P_0$ is a projector onto the fully symmetric manifold;
-$k\in\C_N\p{M_j}$ is a choice of $M_j$ distinct spins; $O_j\p{k}$
-denotes the action of an $M_j$-spin operator on spins $k$ that is
-invariant under permutations of $k$; and $w_j\p{k}$ is a scalar
-coefficient for $O_j\p{k}$.
-
-To simplify the product in \eqref{eq:sym_prod_proj}, we first write
-\begin{align}
-  \prod_{j\in\ZZ_p} \sum_{k\in\C_N\p{M_j}} w_j\p{k} O_j\p{k}
-  = \prod_{j\in\ZZ_p} \f1{M_j!} \sum_{k\in\ZZ_N^{M_j}}
-  w_j\p{k} O_j\p{k},
-  \label{eq:sym_prod_vec}
-\end{align}
-where we extend the definitions of $w_j\p{k}$ and $O_j\p{k}$ to all
-vectors $k\in\ZZ_N^{M_j}$ by defining symmetric rank-$M_j$ tensors
-$w_j$ and $O_j$ (where $O_j$ is an {\it operator-valued} tensor) for
-which $w_j\p{k}=O_j\p{k}=0$ if $k$ contains any repeated symbols.  We
-then collect terms whose operator content is equivalent up to a
-permutation of all spins.  To this end, we classify terms in
-\eqref{eq:sym_prod_vec} by the numbers $g_S$ of indices shared by all
-tensors $w_j,O_j$ with $j\in S\subset\ZZ_p$.  For example, a term of
-the form $w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}$ with distinct
-indices $\p{a,b,c,d,e}\in\C_N\p{5}$ would have
-\begin{align}
-  g_{\set{1}} = 1,
-  &&
-  g_{\set{1,2,3}} = 1,
-  &&
-  g_{\set{1,3}} = 1,
-  &&
-  g_{\set{2,3}} = 2,
-\end{align}
-and $g_S=0$ for all other subsets $S\subset\ZZ_3$.  This index
-assignment can be represented by the Venn diagram
-\begin{align}
-  \diagram{example},
-  \label{eq:venn_diagram}
-\end{align}
-where $g_S$ is determined the number of dots at the intersection of
-circles $j\in S$.  We denote an assignment of distinct indices
-according to choices of $g_S$ for all $S\in\PP\p{\ZZ_p}$ by $g$, where
-$\PP\p{\ZZ_p}$ is the power set (i.e.~set of all subsets) of $\ZZ_p$.
-Given a vector $\m M$ of the tensor ranks $M_j$, we then denote the
-set of all valid index assignments $g$ by $\G\p{\m M}$.  For
-consistency, all valid index assignments $g\in\G\p{\m M}$ must assign
-exactly $M_j$ indices to the rank-$M_j$ tensor $w_j$, i.e.~
-\begin{align}
-  \sum_{S\in\PP\p{\ZZ_p}\,:\,j\in S} g_S
-  = \sum_{R\in\PP\p{\ZZ_p\setminus\set{j}}} g_{\set{j}\cup R}
-  = M_j
-\end{align}
-for all $j\in\ZZ_p$.  In order to make the index assignments
-determined by choices of $g_S$ for all $S\in\PP\p{\ZZ_p}$ unique, we
-also enforce that $g_{\set{}}=0$ for all $g\in\G\p{\m M}$, which is
-equivalent to saying that every dot in the corresponding Venn diagram
-must lie within some circle.  The set of valid index assignments
-$\G\p{\m M}$ is then essentially the set of all Venn diagrams of the
-form in \eqref{eq:venn_diagram} with $p=\rank\p{\m M}$ circles and a
-total of $M_j$ dots within circle $j$.
-
-A classification of terms in \eqref{eq:sym_prod_vec} by index
-assignments $g\in\G\p{\m M}$ allows us to expand
-\begin{align}
-  \prod_{j\in\ZZ_p} \sum_{k\in\ZZ_N^{M_j}} w_j\p{k} O_j\p{k}
-  \EQFS \sum_{g\in\G\p{\m M}} \S\p{g} w\p{g} O\p{g},
-  \label{eq:sym_prod_group_start}
-\end{align}
-where $\EQFS$ denotes equality up to a restriction to the fully
-symmetric manifold, $O\p{g}$ is an operator acquired by assigning
-indices to the tensors $O_j$ in a manner consistent with $g$; $w\p{g}$
-is a scalar acquired by summing over all indices assigned to the
-tensors $w_j$ according to $g$; and $\S\p{g}$ is a symmetry factor
-accounting for the number of equivalent ways to assign indices
-according to $g$.
-
-In order to write out the factors in \eqref{eq:sym_prod_group_start}
-explicitly, we identify the set of values that the distinct indices
-assigned according to $g$ can take:
-\begin{align}
-  \C_N\p{g}
-  \equiv \set{ k \in \bigotimes_{S\in\PP\p{\ZZ_p}} \C_N\p{g_S}
-    : \t{all elements of $k$ are distinct} },
-\end{align}
-where each $k\in\C_N\p{g}$ has the natural decomposition
-$k=\p{k_S:S\in\PP\p{\ZZ_p}}$ with $k_S\in\C_N\p{g_S}$.  We then define
-the restriction of $k\in\C_N\p{g}$ to values associated with $w_j$ and
-$O_j$:
-\begin{align}
-  k_j \equiv \bigcup_{S\in\PP\p{\ZZ_p} : j\in S} k_S
-  = \bigcup_{R\in\PP\p{\ZZ_p\setminus\set{j}}} k_{\set{j}\cup R},
-\end{align}
-where the union of lists $k_S$ denotes a concatenation of those lists.
-These definitions allow us to expand
-\begin{align}
-  w\p{g} \equiv \sum_{k\in\C_N\p{g}} \prod_{j\in\ZZ_p} w_j\p{k_j},
-  &&
-  O\p{g} \equiv \prod_{j\in\ZZ_p}
-  O_j\p{\ell_j}~\t{for some}~\ell\in\C_N\p{g},
-\end{align}
-where the choice of $\ell\in\C_N\p{g}$ does not matter after
-projection onto fully symmetric manifold, and
-\begin{align}
-  \S\p{g} \equiv \sp{\prod_{j\in\ZZ_p} { M_j \choose g_S : j\in S }}
-  \sp{\prod_{S\in\PP\p{\ZZ_p}} \p{g_S!}^\abs{S}}
-  = \prod_{j\in\ZZ_p} M_j!,
-  \label{eq:sym_prod_symmetry}
-\end{align}
-where the first product in \eqref{eq:sym_prod_symmetry} contains
-multinomial coefficients
-\begin{align}
-  { M \choose a_1, a_2, \cdots, a_m }
-  \equiv \f{M!}{a_1!a_2!\cdots a_m!},
-\end{align}
-that account for the number of ways to partition the indices of each
-$w_j,O_j$ into sets of shared indices according to $g_S$ for all
-$j\in S$, and the second product in \eqref{eq:sym_prod_symmetry}
-accounts for the number of ways to permute each set of shared indices
-on every tensor.  Altogether, the symmetry factors $\S\p{g}$ cancel
-out with the $\sim M_j!$ prefactors in \eqref{eq:sym_prod_vec}, so
-\begin{align}
-  \prod_{j\in\ZZ_p} \sum_{k\in\C_N\p{M_j}} w_j\p{k} O_j\p{k}
-  \EQFS \sum_{g\in\G\p{\m M}} w\p{g} O\p{g}.
-  \label{eq:sym_prod_group}
-\end{align}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Two single-body operators}
-\label{sec:PXYP}
-
-Calculating the second-order effective Hamiltonian $H_1^{(2)}$ induced
-on the fully symmetric manifold $\M_0$ by the perturbation $\V_1$ in
-\eqref{eq:perturbations} requires us to simplify a product of the form
-
-\begin{align}
-  \P_0 \sp{\sum_{p,q} v_p w_q X_p Y_q} \P_0
-  \EQFS \diagram{single_body_0} X_1 Y_2
-  + \diagram{single_body_1} X_1 Y_1,
-  \label{eq:PXYP_start}
-\end{align}
-where $p,q$ index individual spins; $X,Y$ are single-spin operators;
-and the coefficients $v_p,w_q$ satisfy \eqref{eq:cond_1} with
-eigenvalue $\Delta\ne0$, which implies that they are mean-zero.
-Writing out
-\begin{align}
-  \diagram{single_body_1}
-  \equiv \sum_p v_p w_p
-  = \v v\c\v w,
-\end{align}
-\begin{align}
-  \diagram{single_body_0}
-  \equiv \sum_{p\ne q} v_p w_q
-  = \sum_{p,q} v_p w_q - \sum_{p=q} v_p w_q
-  = - \v v \c\v w,
-\end{align}
-we thus have
-\begin{align}
-  \sum_{p,q} v_p w_q X_p Y_q
-  \EQFS - \v v\c\v w \p{X_1 Y_2 - X_1 Y_1},
-\end{align}
-This result is also simple to write in terms of collective operators,
-as
-\begin{align}
-  \col{X Y} \EQFS N X_1 Y_1,
-  &&
-  \col{Z} \equiv \sum_p Z_p,
-\end{align}
-\begin{align}
-  \col{X}\,\col{Y}
-  = \sum_{p=q} X_p Y_q + \sum_{p\ne q} X_p Y_q
-  \EQFS \col{XY} + N\p{N-1} X_1 Y_2,
-\end{align}
-so
-\begin{align}
-  \sum_{p,q} v_p w_q X_p Y_q
-  \EQFS - \f{\v v \c\v w}{N\p{N-1}}
-  \p{\col{X}\,\col{Y} - N \col{XY}}.
-\end{align}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Two two-body operators}
-\label{sec:POQP}
-
-Calculating the second-order effective Hamiltonian $H_2^{(2)}$ induced
-on the fully symmetric manifold $\M_0$ by the perturbation $\V_2$ in
-\eqref{eq:perturbations} requires us to simplify
-\begin{multline}
-  \P_0 \sp{\sum_{k<\ell} \sum_{p<q}
-    v_{k\ell} w_{pq} O_{k\ell} Q_{pq}} \P_0 \\
-  \EQFS \diagram{two_body_0} O_{1,2} Q_{3,4}
-  + \diagram{two_body_1} O_{1,2} Q_{1,3}
-  + \diagram{two_body_2} O_{1,2} Q_{1,2},
-  \label{eq:POQP_start}
-\end{multline}
-where $k,\ell,p,q$ index individual spins; $O$ and $Q$ are
-permutationally-symmetric two-spin operators; and the coefficients
-$v_{k\ell},w_{pq}$ satisfy \eqref{eq:cond_2} with eigenvalue
-$\Delta\ne0$, which implies that they are mean-zero.  Defining
-\begin{align}
-  \v{\m m} \equiv \sum_{p<q} m_{pq} \ket{\set{p,q}},
-  &&
-  \v m \equiv \sum_p m_p \ket{p},
-  &&
-  m_p \equiv \sum_q m_{pq},
-\end{align}
-for $m\in\set{v,w}$, we can write out explicitly
-\begin{align}
-  \diagram{two_body_2}
-  \equiv \sum_{p<q} v_{pq} w_{pq} = \v{\m v} \c\v{\m w},
-\end{align}
-and, using the fact that $v_{pp}=w_{pp}=0$,
-\begin{align}
-  \diagram{two_body_1}
-  \equiv \sum_{\substack{k,p,q\\k\ne p\ne q\ne k}} v_{kp} w_{kq}
-  = \sum_{k,p} v_{kp} \sum_{q\ne p} w_{kq}
-  = \sum_{k,p} v_{kp} \p{w_{k} - w_{kp}}
-  = \v v \c \v w - 2 \v{\m v} \c \v{\m w}.
-\end{align}
-Using additionally the fact that $v_{pq},w_{pq}$ are mean-zero, with
-some more work we can work out that
-\begin{align}
-  \diagram{two_body_0}
-  \equiv \sum_{k<\ell} \sum_{\substack{p<q\\p,q\notin\set{k,\ell}}}
-  v_{k\ell} w_{pq} = -\v v \c \v w + \v{\m v} \c \v{\m w}.
-\end{align}
-Altogether, we thus have that
-\begin{multline}
-  \sum_{k<\ell} \sum_{p<q} v_{k\ell} w_{pq} O_{k\ell} Q_{pq} \\
-  \EQFS - \p{\v v \c \v w - \v{\m v} \c \v{\m w}} O_{1,2} Q_{3,4}
-  + \p{\v v\c\v w - 2 \v{\m v} \c \v{\m w}} O_{1,2} Q_{1,3}
-  + \v{\m v} \c \v{\m w}\, O_{1,2} Q_{1,2}.
-  \label{eq:POQP}
-\end{multline}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Three two-body Ising-like operators}
-
-Here we simplify a product of three two-body operators of a particular
-form projected onto the fully symmetric manifold:
-\begin{align}
-  \X \equiv \P_0 \sp{\sum_{k<\ell} \sum_{p<q} \sum_{r<s}
-    u_{k\ell} v_{pq} w_{rs} X_k X_\ell X_p X_q X_r X_s} \P_0,
-  &&
-  X^2 = 1,
-\end{align}
-where $k,\ell,p,q,r,s$ index individual spins; $X$ is a single-spin
-operator; and $u_{k\ell},v_{pq},w_{rs}$ are arbitrary scalar
-coefficients with $m_{pq}=m_{qp}$ and $m_{pp}=0$ for each of
-$m\in\set{u,v,w}$.  Collecting terms according to the number of $X$
-operators that remain after accounting for $X^2=1$, we find that
-\begin{align}
-  \X \EQFS A_6 X^{\otimes 6} + A_4 X^{\otimes 4} + A_2 X^{\otimes 2},
-  \label{eq:triple_multi}
-\end{align}
-with
-\begin{align}
-  A_6 \equiv \diagram{triple_6},
-  &&
-  A_4 \equiv \diagram{triple_4_1} + \diagram{triple_4_0},
-  \label{eq:triple_64}
-\end{align}
-\begin{align}
-  A_2 \equiv \diagram{triple_2_0} + \diagram{triple_2_1}
-  + \diagram{triple_2_2} + \diagram{triple_2_3},
-  \label{eq:triple_2}
-\end{align}
-where a diagram with unlabeled circles implicitly contains a sum over
-all label assignments that result in a distinct diagram, e.g.
-\begin{align}
-  \diagram{triple_6} \equiv \diagram{triple_6_uvw},
-  \label{eq:triple_uvw}
-\end{align}
-\begin{align}
-  \diagram{triple_2_3} \equiv \diagram{triple_2_3_uvw}
-  + \diagram{triple_2_3_vwu} + \diagram{triple_2_3_wuv}.
-  \label{eq:triple_uvw_3}
-\end{align}
-The first diagram in each of $A_6,A_4,A_2$ as written in
-\eqref{eq:triple_64}, \eqref{eq:triple_2} thus has only one assignment
-of labels, as in \eqref{eq:triple_uvw}, while all other unlabeled
-diagrams have three assignments, as in \eqref{eq:triple_uvw_3}.  We
-can write \eqref{eq:triple_multi} in terms of collective operators as
-\begin{align}
-  \X \EQFS
-  \tilde A_6 \col{X}^6 + \tilde A_4 \col{X}^4 + \tilde A_2 \col{X}^2,
-  \label{eq:triple_col}
-\end{align}
-where we have neglected scalar terms without physical consequence, and
-\begin{align}
-  \tilde A_6 \equiv \f{A_6}{N!_6},
-  &&
-  \tilde A_4 \equiv \f{A_4}{N!_4} - 5\p{3N-8} \f{A_6}{N!_6},
-\end{align}
-\begin{align}
-  \tilde A_2 \equiv \f{A_2}{N!_2} - 2\p{3N-4} \f{A_4}{N!_4}
-  + \p{45N^2-210N+184} \f{A_6}{N!_6}.
-\end{align}
-
-\vspace{5cm}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Making use of translational invariance}
+\section{Constructing eigenstates of translationally invariant
+  systems}
 \label{sec:trans_inv}
@@
 to generate states of fixed excitation energy $\Delta$ when the
-corresponding perturbation $\V_2^\Delta$ acts on a fully symmetric
-state $\ket\psi\in\M_0$.  The decomposition of an arbitrary
+corresponding perturbation $\V_2^\Delta$ acts on a permutationally
+symmetric state $\ket\psi\in\M_0$.  The decomposition of an arbitrary
 coefficient vector $\v{\m w}=\sum_\Delta\v{\m w}_\Delta$ into
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Operator products restricted to the permutationally symmetric
+  manifold}
+\label{sec:sym_prod}
+
+\subsection{The general case}
+
+Given a system of $N$ spins, we wish to project a product of
+multi-body operators onto the permutationally symmetric manifold.  For
+$p$ multi-body operators, the projection of such a product takes the
+form
+\begin{align}
+  \P_0 \sp{\prod_{j\in\ZZ_p}
+    \sum_{k\in\C_N\p{M_j}} w_j\p{k} O_j\p{k}} \P_0,
+  \label{eq:sym_prod_proj}
+\end{align}
+where $\P_0$ is a projector onto the permutationally symmetric
+manifold; $k\in\C_N\p{M_j}$ is a choice of $M_j$ distinct spins;
+$O_j\p{k}$ denotes the action of an $M_j$-spin operator on spins $k$
+that is invariant under permutations of $k$; and $w_j\p{k}$ is a
+scalar coefficient for $O_j\p{k}$.
+
+To simplify the product in \eqref{eq:sym_prod_proj}, we first write
+\begin{align}
+  \prod_{j\in\ZZ_p} \sum_{k\in\C_N\p{M_j}} w_j\p{k} O_j\p{k}
+  = \prod_{j\in\ZZ_p} \f1{M_j!} \sum_{k\in\ZZ_N^{M_j}}
+  w_j\p{k} O_j\p{k},
+  \label{eq:sym_prod_vec}
+\end{align}
+where we extend the definitions of $w_j\p{k}$ and $O_j\p{k}$ to all
+vectors $k\in\ZZ_N^{M_j}$ by defining symmetric rank-$M_j$ tensors
+$w_j$ and $O_j$ (where $O_j$ is an {\it operator-valued} tensor) for
+which $w_j\p{k}=O_j\p{k}=0$ if $k$ contains any repeated symbols.  We
+then collect terms whose operator content is equivalent up to a
+permutation of spins.  To this end, we classify terms in
+\eqref{eq:sym_prod_vec} by the numbers $g_S$ of indices shared by all
+tensors $w_j,O_j$ with $j\in S\subset\ZZ_p$.  For example, a term of
+the form $w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}$ with distinct
+indices $\p{a,b,c,d,e}\in\C_N\p{5}$ would have
+\begin{align}
+  g_{\set{1}} = 1,
+  &&
+  g_{\set{1,2,3}} = 1,
+  &&
+  g_{\set{1,3}} = 1,
+  &&
+  g_{\set{2,3}} = 2,
+\end{align}
+and $g_S=0$ for all other subsets $S\subset\ZZ_3$.  This index
+assignment can be represented by the Venn diagram
+\begin{align}
+  \diagram{example},
+  \label{eq:venn_diagram}
+\end{align}
+where $g_S$ is determined the number of dots at the intersection of
+circles $j\in S$.  We denote an assignment of distinct indices
+according to choices of $g_S$ for all $S\in\PP\p{\ZZ_p}$ by $g$, where
+$\PP\p{\ZZ_p}$ is the power set (i.e.~set of all subsets) of $\ZZ_p$.
+Given a vector $\m M$ of the tensor ranks $M_j$, we then denote the
+set of all valid index assignments $g$ by $\G\p{\m M}$.  For
+consistency, all valid index assignments $g\in\G\p{\m M}$ must assign
+exactly $M_j$ indices to the rank-$M_j$ tensor $w_j$, i.e.~
+\begin{align}
+  \sum_{S\in\PP\p{\ZZ_p}\,:\,j\in S} g_S
+  = \sum_{R\in\PP\p{\ZZ_p\setminus\set{j}}} g_{\set{j}\cup R}
+  = M_j
+\end{align}
+for all $j\in\ZZ_p$.  In order to make the index assignments
+determined by choices of $g_S$ for all $S\in\PP\p{\ZZ_p}$ unique, we
+also enforce that $g_{\set{}}=0$ for all $g\in\G\p{\m M}$, which is
+equivalent to saying that every dot in the corresponding Venn diagram
+must lie within some circle.  The set of valid index assignments
+$\G\p{\m M}$ is then essentially the set of all Venn diagrams of the
+form in \eqref{eq:venn_diagram} with $p=\rank\p{\m M}$ circles and a
+total of $M_j$ dots within circle $j$.
+
+A classification of terms in \eqref{eq:sym_prod_vec} by index
+assignments $g\in\G\p{\m M}$ allows us to expand
+\begin{align}
+  \prod_{j\in\ZZ_p} \sum_{k\in\ZZ_N^{M_j}} w_j\p{k} O_j\p{k}
+  \EQFS \sum_{g\in\G\p{\m M}} \S\p{g} w\p{g} O\p{g},
+  \label{eq:sym_prod_group_start}
+\end{align}
+where $\EQFS$ denotes equality up to a restriction to the fully
+symmetric manifold, $O\p{g}$ is an operator acquired by assigning
+indices to the tensors $O_j$ in a manner consistent with $g$; $w\p{g}$
+is a scalar acquired by summing over all indices assigned to the
+tensors $w_j$ according to $g$; and $\S\p{g}$ is a symmetry factor
+accounting for the number of equivalent ways to assign indices
+according to $g$.
+
+In order to write out the factors in \eqref{eq:sym_prod_group_start}
+explicitly, we identify the set of values that the distinct indices
+assigned according to $g$ can take:
+\begin{align}
+  \C_N\p{g}
+  \equiv \set{ k \in \bigotimes_{S\in\PP\p{\ZZ_p}} \C_N\p{g_S}
+    : \t{all elements of $k$ are distinct} },
+\end{align}
+where each $k\in\C_N\p{g}$ has the natural decomposition
+$k=\p{k_S:S\in\PP\p{\ZZ_p}}$ with $k_S\in\C_N\p{g_S}$.  We then define
+the restriction of $k\in\C_N\p{g}$ to values associated with $w_j$ and
+$O_j$:
+\begin{align}
+  k_j \equiv \bigcup_{S\in\PP\p{\ZZ_p} : j\in S} k_S
+  = \bigcup_{R\in\PP\p{\ZZ_p\setminus\set{j}}} k_{\set{j}\cup R},
+\end{align}
+where the union of lists $k_S$ denotes a concatenation of those lists.
+These definitions allow us to expand
+\begin{align}
+  w\p{g} \equiv \sum_{k\in\C_N\p{g}} \prod_{j\in\ZZ_p} w_j\p{k_j},
+  &&
+  O\p{g} \equiv \prod_{j\in\ZZ_p}
+  O_j\p{\ell_j}~\t{for some}~\ell\in\C_N\p{g},
+\end{align}
+where the choice of $\ell\in\C_N\p{g}$ does not matter after
+projection onto permutationally symmetric manifold, and
+\begin{align}
+  \S\p{g} \equiv \sp{\prod_{j\in\ZZ_p} { M_j \choose g_S : j\in S }}
+  \sp{\prod_{S\in\PP\p{\ZZ_p}} \p{g_S!}^\abs{S}}
+  = \prod_{j\in\ZZ_p} M_j!,
+  \label{eq:sym_prod_symmetry}
+\end{align}
+where the first product in \eqref{eq:sym_prod_symmetry} contains
+multinomial coefficients
+\begin{align}
+  { M \choose a_1, a_2, \cdots, a_m }
+  \equiv \f{M!}{a_1!a_2!\cdots a_m!},
+\end{align}
+that account for the number of ways to partition the indices of each
+$w_j,O_j$ into sets of shared indices according to $g_S$ for all
+$j\in S$, and the second product in \eqref{eq:sym_prod_symmetry}
+accounts for the number of ways to permute each set of shared indices
+on every tensor.  Altogether, the symmetry factors $\S\p{g}$ cancel
+out with the $\sim M_j!$ prefactors in \eqref{eq:sym_prod_vec}, so
+\begin{align}
+  \prod_{j\in\ZZ_p} \sum_{k\in\C_N\p{M_j}} w_j\p{k} O_j\p{k}
+  \EQFS \sum_{g\in\G\p{\m M}} w\p{g} O\p{g}.
+  \label{eq:sym_prod_group}
+\end{align}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Two single-body operators}
+\label{sec:PXYP}
+
+Calculating the second-order effective Hamiltonian $H_1^{(2)}$ induced
+on the permutationally symmetric manifold $\M_0$ by the perturbation
+$\V_1$ in \eqref{eq:perturbations} requires us to simplify a product
+of the form
+
+\begin{align}
+  \P_0 \sp{\sum_{p,q} v_p w_q X_p Y_q} \P_0
+  \EQFS \diagram{single_body_0} X_1 Y_2
+  + \diagram{single_body_1} X_1 Y_1,
+  \label{eq:PXYP_start}
+\end{align}
+where $p,q$ index individual spins; $X,Y$ are single-spin operators;
+and the coefficients $v_p,w_q$ satisfy \eqref{eq:cond_1} with
+eigenvalue $\Delta\ne0$, which implies that they are mean-zero.
+Writing out
+\begin{align}
+  \diagram{single_body_1}
+  \equiv \sum_p v_p w_p
+  = \v v\c\v w,
+\end{align}
+\begin{align}
+  \diagram{single_body_0}
+  \equiv \sum_{p\ne q} v_p w_q
+  = \sum_{p,q} v_p w_q - \sum_{p=q} v_p w_q
+  = - \v v \c\v w,
+\end{align}
+we thus have
+\begin{align}
+  \sum_{p,q} v_p w_q X_p Y_q
+  \EQFS - \v v\c\v w \p{X_1 Y_2 - X_1 Y_1},
+\end{align}
+This result is also simple to write in terms of collective operators,
+as
+\begin{align}
+  \col{X Y} \EQFS N X_1 Y_1,
+  &&
+  \col{Z} \equiv \sum_p Z_p,
+\end{align}
+\begin{align}
+  \col{X}\,\col{Y}
+  = \sum_{p=q} X_p Y_q + \sum_{p\ne q} X_p Y_q
+  \EQFS \col{XY} + N\p{N-1} X_1 Y_2,
+\end{align}
+so
+\begin{align}
+  \sum_{p,q} v_p w_q X_p Y_q
+  \EQFS - \f{\v v \c\v w}{N\p{N-1}}
+  \p{\col{X}\,\col{Y} - N \col{XY}}.
+\end{align}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Two two-body operators}
+\label{sec:POQP}
+
+Calculating the second-order effective Hamiltonian $H_2^{(2)}$ induced
+on the permutationally symmetric manifold $\M_0$ by the perturbation
+$\V_2$ in \eqref{eq:perturbations} requires us to simplify
+\begin{multline}
+  \P_0 \sp{\sum_{k<\ell} \sum_{p<q}
+    v_{k\ell} w_{pq} O_{k\ell} Q_{pq}} \P_0 \\
+  \EQFS \diagram{two_body_0} O_{1,2} Q_{3,4}
+  + \diagram{two_body_1} O_{1,2} Q_{1,3}
+  + \diagram{two_body_2} O_{1,2} Q_{1,2},
+  \label{eq:POQP_start}
+\end{multline}
+where $k,\ell,p,q$ index individual spins; $O$ and $Q$ are
+permutationally-symmetric two-spin operators; and the coefficients
+$v_{k\ell},w_{pq}$ satisfy \eqref{eq:cond_2} with eigenvalue
+$\Delta\ne0$, which implies that they are mean-zero.  Defining
+\begin{align}
+  \v{\m m} \equiv \sum_{p<q} m_{pq} \ket{\set{p,q}},
+  &&
+  \v m \equiv \sum_p m_p \ket{p},
+  &&
+  m_p \equiv \sum_q m_{pq},
+\end{align}
+for $m\in\set{v,w}$, we can write out explicitly
+\begin{align}
+  \diagram{two_body_2}
+  \equiv \sum_{p<q} v_{pq} w_{pq} = \v{\m v} \c\v{\m w},
+\end{align}
+and, using the fact that $v_{pp}=w_{pp}=0$,
+\begin{align}
+  \diagram{two_body_1}
+  \equiv \sum_{\substack{k,p,q\\k\ne p\ne q\ne k}} v_{kp} w_{kq}
+  = \sum_{k,p} v_{kp} \sum_{q\ne p} w_{kq}
+  = \sum_{k,p} v_{kp} \p{w_{k} - w_{kp}}
+  = \v v \c \v w - 2 \v{\m v} \c \v{\m w}.
+\end{align}
+Using additionally the fact that $v_{pq},w_{pq}$ are mean-zero, with
+some more work we find that
+\begin{align}
+  \diagram{two_body_0}
+  \equiv \sum_{k<\ell} \sum_{\substack{p<q\\p,q\notin\set{k,\ell}}}
+  v_{k\ell} w_{pq} = -\v v \c \v w + \v{\m v} \c \v{\m w}.
+\end{align}
+Altogether, we thus have that
+\begin{multline}
+  \sum_{k<\ell} \sum_{p<q} v_{k\ell} w_{pq} O_{k\ell} Q_{pq} \\
+  \EQFS - \p{\v v \c \v w - \v{\m v} \c \v{\m w}} O_{1,2} Q_{3,4}
+  + \p{\v v\c\v w - 2 \v{\m v} \c \v{\m w}} O_{1,2} Q_{1,3}
+  + \v{\m v} \c \v{\m w}\, O_{1,2} Q_{1,2}.
+  \label{eq:POQP}
+\end{multline}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Three two-body Ising-like operators}
+
+Here we simplify a product of three two-body operators of a particular
+form projected onto the permutationally symmetric manifold:
+\begin{align}
+  \X \equiv \P_0 \sp{\sum_{k<\ell} \sum_{p<q} \sum_{r<s}
+    u_{k\ell} v_{pq} w_{rs} X_k X_\ell X_p X_q X_r X_s} \P_0,
+  &&
+  X^2 = 1,
+\end{align}
+where $k,\ell,p,q,r,s$ index individual spins; $X$ is a single-spin
+operator; and $u_{k\ell},v_{pq},w_{rs}$ are arbitrary scalar
+coefficients with $m_{pq}=m_{qp}$ and $m_{pp}=0$ for each of
+$m\in\set{u,v,w}$.  Collecting terms according to the number of $X$
+operators that remain after accounting for $X^2=1$, we find that
+\begin{align}
+  \X \EQFS A_6 X^{\otimes 6} + A_4 X^{\otimes 4} + A_2 X^{\otimes 2},
+  \label{eq:triple_multi}
+\end{align}
+with
+\begin{align}
+  A_6 \equiv \diagram{triple_6},
+  &&
+  A_4 \equiv \diagram{triple_4_1} + \diagram{triple_4_0},
+  \label{eq:triple_64}
+\end{align}
+\begin{align}
+  A_2 \equiv \diagram{triple_2_0} + \diagram{triple_2_1}
+  + \diagram{triple_2_2} + \diagram{triple_2_3},
+  \label{eq:triple_2}
+\end{align}
+where a diagram with unlabeled circles implicitly contains a sum over
+all label assignments that result in a distinct diagram, e.g.
+\begin{align}
+  \diagram{triple_6} \equiv \diagram{triple_6_uvw},
+  \label{eq:triple_uvw}
+\end{align}
+\begin{align}
+  \diagram{triple_2_3} \equiv \diagram{triple_2_3_uvw}
+  + \diagram{triple_2_3_vwu} + \diagram{triple_2_3_wuv}.
+  \label{eq:triple_uvw_3}
+\end{align}
+The first diagram in each of $A_6,A_4,A_2$ as written in
+\eqref{eq:triple_64}, \eqref{eq:triple_2} thus has only one assignment
+of labels, as in \eqref{eq:triple_uvw}, while all other unlabeled
+diagrams have three assignments, as in \eqref{eq:triple_uvw_3}.  We
+can write \eqref{eq:triple_multi} in terms of collective operators as
+\begin{align}
+  \X \EQFS
+  \tilde A_6 \col{X}^6 + \tilde A_4 \col{X}^4 + \tilde A_2 \col{X}^2,
+  \label{eq:triple_col}
+\end{align}
+where we have neglected scalar terms without physical consequence, and
+\begin{align}
+  \tilde A_6 \equiv \f{A_6}{N!_6},
+  &&
+  \tilde A_4 \equiv \f{A_4}{N!_4} - 5\p{3N-8} \f{A_6}{N!_6},
+\end{align}
+\begin{align}
+  \tilde A_2 \equiv \f{A_2}{N!_2} - 2\p{3N-4} \f{A_4}{N!_4}
+  + \p{45N^2-210N+184} \f{A_6}{N!_6}.
+\end{align}
+
 % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \section{Collective and multi-local operators in the fully symmetric
-%   manifold}
+% \section{Collective and multi-local operators in the permutationally
+%   symmetric manifold}
 % \label{sec:multi_to_collective}
@@
 % Here we provide a decomposition multi-local operators into sums of
-% collective operators (and vice versa) within the fully symmetric
-% manifold of $N$ spins.  Multi-local operators are operators of the
-% form
+% collective operators (and vice versa) within the permutationally
+% symmetric manifold of $N$ spins.  Multi-local operators are operators
+% of the form
 % \begin{align}
@@
 % order of tensor factors in \eqref{eq:multi_op_set} does not matter
-% after projecting operators onto the fully symmetric manifold, as we
-% always will in this section.  While the decomposition of a collective
-% operator into a sum of multi-local operators is unique, the opposite
-% is not true: there is generally no unique decomposition of a
-% multi-local operator into a sum of collective operators.  We can,
-% however, enforce a canonical decomposition by choosing a set of
-% collective operators that obey the same symmetries as $O_S$, namely
-% invariance under arbitrary permutations of the single-spin operators
-% $O\in S$.
+% after projecting operators onto the permutationally symmetric
+% manifold, as we always will in this section.  While the decomposition
+% of a collective operator into a sum of multi-local operators is
+% unique, the opposite is not true: there is generally no unique
+% decomposition of a multi-local operator into a sum of collective
+% operators.  We can, however, enforce a canonical decomposition by
+% choosing a set of collective operators that obey the same symmetries
+% as $O_S$, namely invariance under arbitrary permutations of the
+% single-spin operators $O\in S$.
 

### 2020-02-04 collective_sun/sun_perturbation.tex
added text (in progress) formalizing diagrams and explaining rules to decompose/manipulate them

@@
 where we extend the definitions of $w_j\p{k}$ and $O_j\p{k}$ to all
-vectors $k\in\ZZ_N^{M_j}$ by defining symmetric rank-$M_j$ tensors
-$w_j$ and $O_j$ (where $O_j$ is an {\it operator-valued} tensor) for
-which $w_j\p{k}=O_j\p{k}=0$ if $k$ contains any repeated symbols.  We
-then collect terms whose operator content is equivalent up to a
-permutation of spins.  To this end, we classify terms in
-\eqref{eq:sym_prod_vec} by the numbers $g_S$ of indices shared by all
-tensors $w_j,O_j$ with $j\in S\subset\ZZ_p$.  For example, a term of
-the form $w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}$ with distinct
-indices $\p{a,b,c,d,e}\in\C_N\p{5}$ would have
+vectors $k\in\ZZ_N^{M_j}$ by defining symmetric dimension-$M_j$
+tensors $w_j$ and $O_j$ (i.e.~tensors with $M_j$ indices, with $O_j$
+an {\it operator-valued} tensor) for which $w_j\p{k}=O_j\p{k}=0$ if
+$k$ contains any repeated symbols.  We then collect terms whose
+operator content is equivalent up to a permutation of spins.  To this
+end, we classify terms in \eqref{eq:sym_prod_vec} by the numbers $g_S$
+of indices shared by all tensors $w_j,O_j$ with $j\in S\subset\ZZ_p$.
+For example, a term of the form
+$w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}$ with distinct indices
+$\p{a,b,c,d,e}\in\C_N\p{5}$ would have
 \begin{align}
@@
 \begin{align}
-  \diagram{example},
+  \diagram{example_triple},
   \label{eq:venn_diagram}
@@
 $\PP\p{\ZZ_p}$ is the power set (i.e.~set of all subsets) of $\ZZ_p$.
-Given a vector $\m M$ of the tensor ranks $M_j$, we then denote the
-set of all valid index assignments $g$ by $\G\p{\m M}$.  For
-consistency, all valid index assignments $g\in\G\p{\m M}$ must assign
-exactly $M_j$ indices to the rank-$M_j$ tensor $w_j$, i.e.~
+Given a vector $\m M$ of the dimensions $M_j$, we then denote the set
+of all valid index assignments $g$ by $\G\p{\m M}$.  For consistency,
+all valid index assignments $g\in\G\p{\m M}$ must assign exactly $M_j$
+indices to the dimension-$M_j$ tensor $w_j$, i.e.~
 \begin{align}
@@
 $\G\p{\m M}$ is then essentially the set of all Venn diagrams of the
-form in \eqref{eq:venn_diagram} with $p=\rank\p{\m M}$ circles and a
+form in \eqref{eq:venn_diagram} with $p=\dim\p{\m M}$ circles and a
 total of $M_j$ dots within circle $j$.
@@
     : \t{all elements of $k$ are distinct} },
+  \label{eq:index_values}
 \end{align}
 where each $k\in\C_N\p{g}$ has the natural decomposition
-$k=\p{k_S:S\in\PP\p{\ZZ_p}}$ with $k_S\in\C_N\p{g_S}$.  We then define
-the restriction of $k\in\C_N\p{g}$ to values associated with $w_j$ and
-$O_j$:
+$k=\p{k_S:S\in\PP\p{\ZZ_p}}$ with $k_S\in\C_N\p{g_S}$.  In words,
+$\C_N\p{g}$ consists of all the ways to choose, for each subset $S$ of
+the tensors $\set{w_j}$, a set $k_S$ of $g_S$ distinct spins, with the
+restriction that all $k_S$ are orthogonal: if $z\in k_S$, then
+$z\notin k_R$ for $R\ne S$.  Having defined the set $\C_N\p{g}$ of
+index values, we define the restriction of $k\in\C_N\p{g}$ to values
+associated with $w_j$ and $O_j$:
 \begin{align}
@@
   O_j\p{\ell_j}~\t{for some}~\ell\in\C_N\p{g},
+  \label{eq:diagram_factors}
 \end{align}
@@
   { M \choose a_1, a_2, \cdots, a_m }
-  \equiv \f{M!}{a_1!a_2!\cdots a_m!},
+  \equiv \f{M!}{a_1!a_2!\cdots a_m!}
 \end{align}
@@
 \end{align}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Formalizing diagrams}
+\label{sec:diagrams}
+
+Diagrams such as \eqref{eq:venn_diagram} will play a major role in the
+calculations in this section.  Here, we define diagrams as formal
+objects that we can use in mathematical expressions.  Formalizing
+diagrams will enable us to perform necessary calculations that are
+otherwise intractable.
+
+The simplest diagrams have $g_S$ dots (``$\bullet$'') at the
+intersection of the circles $j\in S$, and are simply equal to the
+associated scalar $w\p{g}$ defined in \eqref{eq:diagram_factors}, e.g.
+\begin{align}
+  \diagram{example_triple}
+  \equiv \sum_{\substack{a,b,c\in\ZZ_N\\\p{d,e}\in\C_N\p{2}\\
+      \abs{\set{a,b,c,d,e}}=5}}
+  w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}.
+  \label{eq:diagram_definition}
+\end{align}
+Additionally, we will define diagrams in which dots may be replaced by
+colored crosses (``$\bm\times$'') or colored pluses (``$\bm+$'').
+Whereas a diagram with $\abs{g}$ dots defines a scalar that takes
+$\O\p{N^\abs{g}}$ time to compute, diagrams with crosses and pluses
+will be used to compute the same scalar with a reduced computational
+cost.
+
+In a diagram with dots, crosses, and pluses, the values of indices
+associated with dots are summed as if the diagram had only dots,
+treating all other indices as independent and fixed.  Crosses in a
+diagram then correspond to indices that are summed over all of $\ZZ_N$
+without additional constraints, e.g.
+\begin{align}
+  \diagram{example_triple_xx_2} \equiv
+  \sum_{a,c\in\ZZ_N}
+  \sum_{\substack{b\in\ZZ_N\\\p{d,e}\in\C_N\p{2}\\
+      \abs{\set{b,d,e}}=3}}
+  w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}.
+\end{align}
+If two crosses have the same color, then the corresponding indices
+must be equal:
+\begin{align}
+  \diagram{example_triple_xx_1} \equiv
+  \sum_{c\in\ZZ_N} \sum_{\substack{a,b,d\in\ZZ_N\\\abs{\set{a,b,d}}=3}}
+  w_1\p{a,b,c} w_2\p{b,d,c} w_3\p{b,c,d,c}.
+\end{align}
+Crosses thus correspond to indices that are entirely independent of
+all other indices.  Pluses, meanwhile, correspond to indices that are
+{\it maximally dependent} on the dots: ``plus indices'' are summed
+{\it only} over the values of ``dot indices''.  As with crosses, the
+same color on two pluses indicates that the corresponding indices are
+equal, e.g.
+\begin{align}
+  \diagram{example_triple_pp_1}
+  \equiv \sum_{\substack{a\in\ZZ_N\\\p{d,e}\in\C_N\p{2}\\
+      \abs{\set{a,d,e}}=3}} \sum_{b\in\set{a,d,e}}
+  w_1\p{a,b,b} w_2\p{b,d,e} w_3\p{b,b,d,e}.
+\end{align}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Decomposing diagrams to reduce computational costs}
+\label{sec:costs}
+
+The diagrams generated by a product of multi-body operators in
+\eqref{eq:sym_prod_group} generally have only dots, without crosses or
+pluses.  Such a diagram with $\abs{g}$ dots nominally takes
+$\O\p{N^{\abs{g}}}$ time to compute.  In practice, this cost can be
+prohibitive for performing numerical simulations of systems of any
+appreciable size.  Here, we discuss the decomposition of the diagrams
+that appear in \eqref{eq:sym_prod_group} into diagrams with reduced
+computational cost.  As an example, we can decompose
+\begin{align}
+  \diagram{example_triple}
+  = \diagram{example_triple_x_top} - \diagram{example_triple_p_top},
+  \label{eq:example_decomposition}
+\end{align}
+where
+\begin{align}
+  \diagram{example_triple}
+  \equiv \sum_{\substack{a,b,c\in\ZZ_N\\\p{d,e}\in\C_N\p{2}\\
+      \abs{\set{a,b,c,d,e}}=5}}
+  w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e},
+\end{align}
+\begin{align}
+  \diagram{example_triple_p_top}
+  \equiv \sum_{\substack{b,c\in\ZZ_N\\\p{d,e}\in\C_N\p{2}\\
+      \abs{\set{b,c,d,e}}=4\\a\in\set{b,c,d,e}}}
+  w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e},
+\end{align}
+\begin{align}
+  \diagram{example_triple_x_top}
+  \equiv \sum_{\substack{b,c\in\ZZ_N\\\p{d,e}\in\C_N\p{2}\\
+      \abs{\set{b,c,d,e}}=4}}
+  w_1\p{\times,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e},
+  \label{eq:example_triple_x_top}
+\end{align}
+with
+\begin{align}
+  w_1\p{\times,b,c} \equiv \sum_{a\in\ZZ_N} w_1\p{a,b,c},
+\end{align}
+The decomposition in \eqref{eq:example_decomposition} essentially
+breaks up a sum over $a\in\ZZ_N\setminus\set{b,c,d,e}$ into a
+difference of sums over $a\in\ZZ_N$ and $a\in\set{b,c,d,e}$.  The sum
+over $a\in\ZZ_N$ to compute $w_1\p{\times,b,c}$ has $\O\p{N}$ cost,
+and can be performed prior to computing the diagram in
+\eqref{eq:example_triple_x_top}.  The decomposition in
+\eqref{eq:example_decomposition} thereby reduces the calculation of
+one $\O\p{N^5}$ diagram to the calculation of two $\O\p{N^4}$
+diagrams.
+
+In general, in a diagram with only dots and crosses, any dot may be
+decomposed similarly to \eqref{eq:example_decomposition} into a cross
+and a plus (with a new color).  The only catch in making such a
+replacement is that one must manually keep track of combinatorial
+factors: if the replaced dot was in a region of the diagram with $D$
+dots, then the new diagrams pick up a factor of $1/D$, e.g.
+\begin{align}
+  \diagram{example_double}
+  = \f13 \diagram{example_double_x}
+  - \f13 \diagram{example_double_p}.
+\end{align}
+This factor can be derived by first treating the values $\bar R$ of
+dots {\it outside} the region of interest fixed, and replacing the sum
+over $\C_N\p{D}\setminus\bar R$ for the dots {\it within} the region
+by a sum over $\ZZ_N^D\setminus\bar R$.  This replacement picks up a
+factor of $1/D!$ to account for the number of ways to permute $D$
+values.  After setting aside one dot for replacement by a cross and a
+plus, the sum over $\ZZ_N^{D-1}\setminus\bar R$ for the remaining dots
+can be changed back to a sum over $\C_N\p{D-1}\setminus\bar R$,
+picking up a factor of $\p{D-1}!$.  The overall factor picked up
+throughout this procedure is therefore $\p{D-1}!/D!=1/D$.
+
+While it is possible to derive rules for decomposing dots in a diagram
+with pluses, it is easier to instead eliminate pluses from a diagram
+altogether.
+
+\vspace{5cm}
 
@@
 \end{align}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Reducing computational costs}
+
+The results in \eqref{eq:triple_multi} and \eqref{eq:triple_col}
+suffer from the fact that the coefficients $A_6$, $A_4$, and $A_2$
+respectively take $\O\p{N^6}$, $\O\p{N^5}$, and $\O\p{N^4}$ time to
+compute in general.  Here, we reduce the computational cost of these
+coefficients.
+
+We begin with the (only) diagram for $A_6$:
+\begin{align}
+  \diagram{triple_6}
+  &= \sum_{\substack{\p{k,\ell},\p{p,q},\p{r,s}\in\C_N\p{2}\\
+      \abs{\set{k,\ell,p,q,r,s}}=6}}
+  u_{k\ell} v_{pq} w_{rs} \\
+  &= \f12 \sum_{k,\ell\in\ZZ_N} u_{k\ell}
+  \  \f12 \sum_{p,q\in\ZZ_N\setminus\set{k,\ell}} v_{pq}
+  \  \f12 \sum_{r,s\in\ZZ_N\setminus\set{k,\ell,p,q}} w_{rs},
+\end{align}
+where we first simplify
+\begin{align}
+  \sum_{r,s\in\ZZ_N\setminus\set{k,\ell,p,q}} w_{rs}
+  = \col{w}_{k\ell pq}
+  \equiv \col{w} - 2 \sum_{r\in\set{k,\ell,p,q}} w_r
+  + \sum_{r,s\in\set{k,\ell,p,q}} w_{rs},
+\end{align}
+with
+\begin{align}
+  \col{w} \equiv \sum_{r,s} w_{rs},
+  &&
+  w_r \equiv \sum_{s\in\ZZ_N} w_{rs}.
+\end{align}
+In turn,
+\begin{align}
+  \sum_{p,q\in\ZZ_N\setminus\set{k,\ell}} v_{pq} \, \col{w}_{k\ell pq}
+  &=
+\end{align}
+
+\vspace{5cm}
 

### 2020-02-04 collective_sun/scripts/make_diagrams.py

@@ -70,24 +70,36 @@ region_mids["ac"] = rotate("b") @ region_mids["bc"]
 ##################################################
 # define constructor for diagram
 
-def place_dots(region, num, region_mids = region_mids, sep = radius/3,
-               marker = "k.", markersize = 5):
-    point = region_mids[region]
+def place_dots(region, num, markers = None, region_mids = region_mids,
+               sep = radius/3, markersize = 3):
+    if markers is None:
+        markers = "ko" * num
+
+    vec_list = []
+
     if num == 1:
-        plt.plot(*point, marker, markersize = markersize)
+        vec_list = [ np.zeros(2) ]
 
     elif num == 2:
         sep_vec = np.array([sep,0])
-        point_1 = point - sep_vec/2
-        point_2 = point + sep_vec/2
-        plt.plot(*point_1, marker, markersize = markersize)
-        plt.plot(*point_2, marker, markersize = markersize)
+        vec_list = [ -sep_vec/2, +sep_vec/2 ]
 
-    elif num > 2:
+    elif num == 3:
+        sep_vec = np.array([0,sep])
+        vec_list = [ rotate(jj) @ sep_vec / np.sqrt(3) for jj in range(3) ]
+
+    elif num > 3:
         print("capability to place more than two dots not implemented")
         exit()
 
-def make_diagram(region_dots, figsize = figsize, pad = radius/20):
+    point = region_mids[region]
+    marker_list = [ markers[2*jj:2*jj+2] for jj in range(num) ]
+    for vec, marker in zip(vec_list, marker_list):
+        pos = point + vec
+        size = markersize + ( 1 if marker[1] == "+" else 0 )
+        plt.plot(*pos, marker, markersize = size)
+
+def make_triple_diagram(region_dots, markers = {}, figsize = figsize, pad = radius/20):
     fig = plt.figure(figsize = figsize)
     plt.axes().set_aspect("equal")
 
@@ -117,14 +129,14 @@ def make_diagram(region_dots, figsize = figsize, pad = radius/20):
 
[…]

### 2020-02-04 collective_sun/sun_perturbation.tex
modified the definition of diagrams to simply the rules used to decompose them

--- 2020-02-03 collective_sun/sun_perturbation.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Formalizing diagrams and reducing computational costs}
+\label{sec:diagrams}
+
+Diagrams such as \eqref{eq:venn_diagram} will play a major role in the
+calculations in this section.  Here, we first define diagrams as
+formal objects that we can use in mathematical expressions.
+Formalizing diagrams will enable us to perform necessary calculations
+that are otherwise intractable.  We then discuss rules for decomposing
+these diagrams in order to reduce the computational cost of their
+numerical evaluation.
+
+The simplest diagrams have $g_S$ dots (``$\bullet$'') at the
+intersection of the circles $j\in S$, and are proportional to the
+associated scalar $w\p{g}$ defined in \eqref{eq:diagram_factors}, e.g.
+\begin{align}
+  \diagram{example_triple}
+  \equiv \sp{\prod_{S\in\PP\p{\ZZ_p}} g_S!} w\p{g}
+  = \sum_{\substack{a,b,c,d,e\in\ZZ_N\\
+      \abs{\set{a,b,c,d,e}}=5}}
+  w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}.
+  \label{eq:diagram_definition}
+\end{align}
+The proportionality factor accounts for the number of ways to permute
+the indices within each region $S\in\PP\p{\ZZ_p}$.  Including this
+symmetry factor in the definition of a diagram is a matter of choice
+that will ultimately make it easier to work with diagrams.  A diagram
+such at that in \eqref{eq:diagram_definition} with $\abs{g}$ dots
+nominally takes $\O\p{N^{\abs{g}}}$ time to compute.  In practice,
+this computational cost can be prohibitive for performing numerical
+simulations of systems of any appreciable size.  If possible, we
+therefore wish to simplify diagrams in order to reduce their
+computational complexity.  To this end, we first need to define simple
+modifications of diagrams in which dots may be replaced by empty dots
+(``$\circ$'') or crosses (``$\bm\times$'').
+
+In any diagram (with filled dots, empty dots, or crosses), the values
+of indices associated with filled dots are always summed as if the
+diagram had only dots, without regard for all other indices.  Empty
+dots in a diagram then correspond to indices that are summed over all
+of $\ZZ_N$ independently and without additional constraints, e.g.
+\begin{align}
+  \diagram{example_triple_oo_2} \equiv
+  \sum_{a,c\in\ZZ_N}
+  \sum_{\substack{b,d,e\in\ZZ_N\\\abs{\set{b,d,e}}=3}}
+  w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}.
+\end{align}
+Crosses, meanwhile, correspond to indices that are {\it maximally
+  dependent} on filled dots: ``cross indices'' are summed {\it only}
+over the values of ``filled dot indices'', e.g.
+\begin{align}
+  \diagram{example_triple_xx_1}
+  \equiv \sum_{\substack{a,d,e\in\ZZ_N\\\abs{\set{a,d,e}}=3}}
+  \sum_{b,c\in\set{a,d,e}}
+  w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}.
+\end{align}
+Empty dots and crosses allow us to decompose diagrams into sums of
+different diagrams that (may) require less time to evaluate.  We will
+essentially use two tricks to decompose diagrams: eliminating filled
+dots, and eliminating crosses.
+
+Given a diagram containing only dots (filled or empty), we can
+decompose any filled dot into an empty dot and a cross.  For example
+\begin{align}
+  \diagram{example_triple}
+  = \diagram{example_triple_o_top} - \diagram{example_triple_x_top},
+  \label{eq:example_decomposition}
+\end{align}
+where
+\begin{align}
+  \diagram{example_triple}
+  \equiv \sum_{\substack{a,b,c,d,e\in\ZZ_N\\\abs{\set{a,b,c,d,e}}=5}}
+  w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e},
+\end{align}
+\begin{align}
+  \diagram{example_triple_x_top}
+  \equiv \sum_{\substack{b,c,d,e\in\ZZ_N\\\abs{\set{b,c,d,e}}=4}}
+  \sum_{a\in\set{b,c,d,e}}
+  w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e},
+\end{align}
+\begin{align}
+  \diagram{example_triple_o_top}
+  \equiv \sum_{\substack{b,c,d,e\in\ZZ_N\\\abs{\set{b,c,d,e}}=4}}
+  w_1\p{\circ,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e},
+  \label{eq:example_triple_x_top}
+\end{align}
+with
+\begin{align}
+  w_1\p{\circ,b,c} \equiv \sum_{a\in\ZZ_N} w_1\p{a,b,c},
+\end{align}
+The decomposition in \eqref{eq:example_decomposition} essentially
+breaks up a sum over $a\in\ZZ_N\setminus\set{b,c,d,e}$ into a
+difference of sums over $a\in\ZZ_N$ and $a\in\set{b,c,d,e}$.  The sum
+over $a\in\ZZ_N$ to compute $w_1\p{\circ,b,c}$ has $\O\p{N^3}$ cost,
+and can be performed prior to computing the diagram in
+\eqref{eq:example_triple_x_top}.  The decomposition in
+\eqref{eq:example_decomposition} thereby reduces the calculation of
+one $\O\p{N^5}$ diagram to the calculation of two $\O\p{N^4}$
+diagrams.  A similar reduction by a factor of $\sim N$ can be achieved
+by decomposing any dot in a region $S\in\PP\p{\ZZ_p}$ addressing only
+one tensor, i.e.~with $\abs{S}=1$.
+
+While it is possible, in principle, to derive rules for decomposing
+filled dots in a diagram that contains crosses, it is much easier to
+simply eliminate crosses from a diagram altogether.  To eliminate a
+cross from a diagram, we first note that the corresponding index need
+not sum over the values of indices in any tensor that contains that
+index; that is (hint: pay attention to the second sum on each line),
+\begin{align}
+  \diagram{example_triple_x_top_elim}
+  &\equiv \sum_{\substack{b,c,d,e,f\in\ZZ_N\\\abs{\set{b,c,d,e,f}}=5}}
+  \sum_{a\in\set{b,c,d,e,f}}
+  w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,f} \\
+  &= \sum_{\substack{b,c,d,e,f\in\ZZ_N\\\abs{\set{b,c,d,e,f}}=5}}
+  \sum_{a\in\set{d,e,f}}
+  w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,f}.
+  \label{eq:x_elim}
+\end{align}
+For each remaining index addressed by a cross, meanwhile, the cross
+simply forces the corresponding indices to be equal, so
+\begin{align}
+  \diagram{example_triple_x_top_elim}
+  = 2 \diagram{example_triple_x_lft_elim}
+  + \diagram{example_triple_x_rht_elim},
+\end{align}
+where the factor of 2 appears simply to account for both of
+$a\in\set{d,e}$ in \eqref{eq:x_elim}.  The rules derived here for
+replacing filled dots and crosses will be used in the following
+sections to simplify diagrams and minimize the computational cost of
+their evaluation.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Two single-body operators}
@@
 of the form
-
 \begin{align}
   \P_0 \sp{\sum_{p,q} v_p w_q X_p Y_q} \P_0
-  \EQFS \diagram{single_body_0} X_1 Y_2
+  \EQFS \diagram{single_body_2} X_1 Y_2
   + \diagram{single_body_1} X_1 Y_1,
@@
 eigenvalue $\Delta\ne0$, which implies that they are mean-zero.
-Writing out
+Defining $\v m \equiv \sum_p m_p \ket{p}$ for $m\in\set{v,w}$, we can
+simplify
 \begin{align}
@@
 \end{align}
-\begin{align}
-  \diagram{single_body_0}
-  \equiv \sum_{p\ne q} v_p w_q
-  = \sum_{p,q} v_p w_q - \sum_{p=q} v_p w_q
+and
+\begin{align}
+  \diagram{single_body_2}
+  = \diagram{single_body_2_o} - \diagram{single_body_2_x}
+  = - \diagram{single_body_1}
   = - \v v \c\v w,
 \end{align}
-we thus have
+so
 \begin{align}
@@
 \end{align}
-This result is also simple to write in terms of collective operators,
-as
+In order to write this result in terms of collective operators
+$\col{Z} \equiv \sum_p Z_p$, we expand
 \begin{align}
@@
   &&
-  \col{Z} \equiv \sum_p Z_p,
-\end{align}
-\begin{align}
   \col{X}\,\col{Y}
@@
 \end{align}
-so
+which implies that
 \begin{align}
@@
 on the permutationally symmetric manifold $\M_0$ by the perturbation
-$\V_2$ in \eqref{eq:perturbations} requires us to simplify
+$\V_2$ in \eqref{eq:perturbations} requires us to simplify a product
+of the form
 \begin{multline}
@@
     v_{k\ell} w_{pq} O_{k\ell} Q_{pq}} \P_0 \\
-  \EQFS \diagram{two_body_0} O_{1,2} Q_{3,4}
-  + \diagram{two_body_1} O_{1,2} Q_{1,3}
-  + \diagram{two_body_2} O_{1,2} Q_{1,2},
+  \EQFS \f14 \diagram{two_body_4} O_{1,2} Q_{3,4}
+  + \diagram{two_body_3} O_{1,2} Q_{1,3}
+  + \f12 \diagram{two_body_2} O_{1,2} Q_{1,2},
   \label{eq:POQP_start}
@@
 where $k,\ell,p,q$ index individual spins; $O$ and $Q$ are
-permutationally-symmetric two-spin operators; and the coefficients
+permutationally-symmetric two-spin operators; the coefficients
 $v_{k\ell},w_{pq}$ satisfy \eqref{eq:cond_2} with eigenvalue
-$\Delta\ne0$, which implies that they are mean-zero.  Defining
+$\Delta\ne0$, which implies that they are mean-zero; and the factors
+of $1/4$ and $1/2$ are included to cancel out the symmetry factors
+that appear in the definition of a diagram.  Defining
 \begin{align}
@@
 \end{align}
-for $m\in\set{v,w}$, we can write out explicitly
-\begin{align}
-  \diagram{two_body_2}
-  \equiv \sum_{p<q} v_{pq} w_{pq} = \v{\m v} \c\v{\m w},
-\end{align}
-and, using the fact that $v_{pp}=w_{pp}=0$,
-\begin{align}
-  \diagram{two_body_1}
-  \equiv \sum_{\substack{k,p,q\\k\ne p\ne q\ne k}} v_{kp} w_{kq}
-  = \sum_{k,p} v_{kp} \sum_{q\ne p} w_{kq}
-  = \sum_{k,p} v_{kp} \p{w_{k} - w_{kp}}
-  = \v v \c \v w - 2 \v{\m v} \c \v{\m w}.
-\end{align}
-Using additionally the fact that $v_{pq},w_{pq}$ are mean-zero, with
-some more work we find that
-\begin{align}
-  \diagram{two_body_0}
-  \equiv \sum_{k<\ell} \sum_{\substack{p<q\\p,q\notin\set{k,\ell}}}
-  v_{k\ell} w_{pq} = -\v v \c \v w + \v{\m v} \c \v{\m w}.
+for $m\in\set{v,w}$, we can write
+\begin{align}
+  \diagram{two_body_2} = \sum_{p,q\in\ZZ_N} v_{pq} w_{pq}
+  = 2 \sum_{p,q\in\C_N\p{2}} v_{pq} w_{pq}
+  = \v{\m v} \c\v{\m w},
+\end{align}
+and simplify
+\begin{align}
+  \diagram{two_body_3}
+  &= \diagram{two_body_3_o} - \diagram{two_body_3_x} \\
+  &= \diagram{two_body_3_oo} - \diagram{two_body_3_ox}
+  - \diagram{two_body_2} \label{eq:mid_zero} \\
+  &= \v v \c \v w - 2 \v{\m v} \c \v{\m w},
+\end{align}
+where the middle diagram in \eqref{eq:mid_zero} contains an empty sum
+for the cross, and is therefore equal to zero.  Similarly neglecting
+diagrams with empty sums and additionally using the fact that
+$v_{pq},w_{pq}$ are mean-zero, we can work out
+\begin{align}
+  \diagram{two_body_4}
+  &= \diagram{two_body_4_o} - \diagram{two_body_4_x} \\
+  &= \sp{\diagram{two_body_4_oo} - \diagram{two_body_4_ox}}
+  - 2 \sp{\diagram{two_body_3}} \\
+  &= \sp{-2 \diagram{two_body_3_o}}
+  - 2 \sp{\diagram{two_body_3_oo} - \diagram{two_body_2}} \\
+  &= -4 \diagram{two_body_3_oo} + 2 \diagram{two_body_2} \\
+  &= -4 \v v\c\v w + 4 \v{\m w}\c\v{\m w}.
 \end{align}

### 2020-02-04 collective_sun/scripts/make_diagrams.py

@@ -7,7 +7,7 @@ import matplotlib.patches as patches
 figsize = 2 * (0.7,)
 
 radius = 1
-dist = 0.6
+dist = 0.65
 
 font_size = 11
 
@@ -70,36 +70,53 @@ region_mids["ac"] = rotate("b") @ region_mids["bc"]
 ##################################################
 # define constructor for diagram
 
-def place_dots(region, num, markers = None, region_mids = region_mids,
-               sep = radius/3, markersize = 3):
+def place_dots(region, num, markers = None, dot_sep = None,
+               region_mids = region_mids, markersize = 3):
+    if dot_sep == None:
+        dot_sep = 1
+    dot_sep *= radius / 3
     if markers is None:
-        markers = "ko" * num
+        markers = "." * num
+    if len(markers) < num:
+        markers += "." * (num-len(markers))
 
-    vec_list = []
+    vecs = []
 
     if num == 1:
-        vec_list = [ np.zeros(2) ]
+        vecs = [ np.zeros(2) ]
 
     elif num == 2:
-        sep_vec = np.array([sep,0])
-        vec_list = [ -sep_vec/2, +sep_vec/2 ]
+        sep_vec = np.array([dot_sep,0])
+        vecs = [ -sep_vec/2, +sep_vec/2 ]
 
     elif num == 3:
-        sep_vec = np.array([0,sep])
-        vec_list = [ rotate(jj) @ sep_vec / np.sqrt(3) for jj in range(3) ]
+        sep_vec = np.array([0,dot_sep])
+        vecs = [ rotate(jj) @ sep_vec / np.sqrt(3) for jj in range(3) ]
 
     elif num > 3:
         print("capability to place more than two dots not implemented")
         exit()
 
     point = region_mids[region]
-    marker_list = [ markers[2*jj:2*jj+2] for jj in range(num) ]
-    for vec, marker in zip(vec_list, marker_list):
-        pos = point + vec
-        size = markersize + ( 1 if marker[1] == "+" else 0 )
-        plt.plot(*pos, marker, markersize = size)
[…]

### 2020-02-04 soc_squeezing/collective_paper/collective_spin.tex
pushing final, published versions of the collective spin paper

@@
 \item simulation results can only be trusted up to the time at which
-  the initial value of operators $\hat\S_{\v m}$ with weight
+  the initial values of operators $\hat\S_{\v m}$ with weight
   $w\p{\v m}\ge W$ have a non-negligible contribution to expectation
@@
 \end{align}
-with $\delta_{\v m\v n}=1$ if $\v m=\v n$ and 0 otherwise.  For
+with $\delta_{\v m\v n}=1$ if $\v m=\v n$ and zero otherwise.  For
 sufficiently short times, we can truncate the series in
@@
 \end{align}
-We refer Eq.~\eqref{eq:TST} as the truncated short-time (TST)
+We refer to Eq.~\eqref{eq:TST} as the truncated short-time (TST)
 expansion of Heisenberg operators.  Note that when computing an
@@
 physical value of a correlator, i.e.~with
-$\abs{\bk{\hat S_{\v n}\p{t}}}\lesssim S^{\abs{\v n}}$.  Finally, the
-TST expansion is essentially perturbative in the time $t$, which
-implies that its validity as a formal expansion eventually breaks
-down.  Precisely characterizing the implications of these last two
-considerations for the TST expansion requires additional analysis that
-we defer to future work.  An investigation of connections between the
-TST expansion and past work related to the Mori
-formalism\cite{mori1965continuedfraction, annett1994recursive}, for
-example, might answer questions about the breakdown and convergence of
-the TST expansion.  As we show from benchmarks of the TST expansion in
-Section \ref{sec:squeezing}, however, a detailed understanding of
-breakdown is not necessary to diagnose the breakdown time
-$t_{\t{break}}^{(M)}$ beyond which the TST expansion yields inaccurate
-results.  Empirically, we find that going beyond order $M\approx35$
-yields no significant gains in all cases considered in this work.
+$\abs{\bk{\hat S_{\v n}\p{t}}}\lesssim S^{\abs{\v n}}$ (where
+$S\equiv N/2$).  Finally, the TST expansion is essentially
+perturbative in the time $t$, which implies that its validity as a
+formal expansion eventually breaks down.  Precisely characterizing the
+implications of these last two considerations for the TST expansion
+requires additional analysis that we defer to future work.  An
+investigation of connections between the TST expansion and past work
+related to the Mori formalism\cite{mori1965continuedfraction,
+  annett1994recursive}, for example, might answer questions about the
+breakdown and convergence of the TST expansion.  As we show from
+benchmarks of the TST expansion in Section \ref{sec:squeezing},
+however, a detailed understanding of breakdown is not necessary to
+diagnose the breakdown time $t_{\t{break}}^{(M)}$ beyond which the TST
+expansion yields inaccurate results.  Empirically, we find that going
+beyond order $M\approx35$ yields no significant gains in all cases
+considered in this work.
 
@@
   garcia-mata2018chaos}.  In typical scenarios, such squared
-commutators initially vanish by construction through to a choice of
+commutators initially vanish by construction through a choice of
 spatially separated operators.  Collective spin systems, however, have
@@
 example for the type of behavior that is accessible at short times
-with the TST expansion.  These examples are straightforward extend to
-equilibrium settings and spin-boson systems.
+with the TST expansion.  These examples are straightforward to extend
+to equilibrium settings and spin-boson systems.
 
@@
   Office of Scientific Research (AFOSR) grant FA9550-18-1-0319; the
-  AFOSR Multidisciplinary University Research Initiative (MURI) grant;
-  the Defense Advanced Research Projects Agency (DARPA) and Army
-  Research Office (ARO) grant W911NF-16-1-0576 and W911NF-19-1-0210;
-  the National Science Foundation (NSF) grant PHY-1820885; JILA-NSF
-  grant PFC-173400; and the National Institute of Standards and
-  Technology (NIST).
+  Defense Advanced Research Projects Agency (DARPA) and Army Research
+  Office (ARO) grant W911NF-16-1-0576 and W911NF-19-1-0210; the
+  National Science Foundation (NSF) grant PHY-1820885; JILA-NSF grant
+  PFC-1734006; and the National Institute of Standards and Technology
+  (NIST).
 \end{acknowledgments}
@@
 \end{align}
-We now need rearrange the operators in $C_{mn;\mu}$ into a standard
+We now need to rearrange the operators in $C_{mn;\mu}$ into a standard
 order, which means pushing all $S_\z$ operators to the right and, for
@@
 \end{align}
-which satisfy same commutation relations as the standard Pauli
+which satisfy the same commutation relations as the standard Pauli
 operators.  These spin operators can be more compactly defined in the
@@
 loss cannot be constructed from spin operators, which are generally
-bilinear in particle creation/annihilation operators.  When working on
-the level of a spin model, therefore, we can at best only {\it
+bilinear in particle creation or annihilation operators.  When working
+on the level of a spin model, therefore, we can at best only {\it
   emulate} the effect of particle loss by some indirect means.
@@
 $10^3$ trajectories per data point.  While the OAT and TAT models
-produce more squeezing than TNT model with weak decoherence, this
+produce more squeezing than the TNT model with weak decoherence, this
 squeezing falls off faster with an increasing decoherence rate

### 2020-02-05 collective_sun/sun_perturbation.tex
fixed minor error with diagram prefactors
changed the definition of an unlabeled diagram

@@
   = \diagram{single_body_2_o} - \diagram{single_body_2_x}
-  = - \diagram{single_body_1}
+  = \diagram{single_body_2_oo} - \diagram{single_body_1}
   = - \v v \c\v w,
@@
 \begin{align}
-  A_6 \equiv \diagram{triple_6},
-  &&
-  A_4 \equiv \diagram{triple_4_1} + \diagram{triple_4_0},
+  A_6 \equiv \f18 \diagram{triple_6},
+  &&
+  A_4 \equiv \diagram{triple_4_1} + \f32 \diagram{triple_4_0},
   \label{eq:triple_64}
@@
 \begin{align}
-  A_2 \equiv \diagram{triple_2_0} + \diagram{triple_2_1}
-  + \diagram{triple_2_2} + \diagram{triple_2_3},
+  A_2 \equiv \f12 \diagram{triple_2_0} + 3 \diagram{triple_2_1}
+  + \f34 \diagram{triple_2_2} + 3 \diagram{triple_2_3},
   \label{eq:triple_2}
 \end{align}
-where a diagram with unlabeled circles implicitly contains a sum over
-all label assignments that result in a distinct diagram, e.g.
+where a diagram with unlabeled circles denotes an average over all
+label assignments that result in a distinct diagram, e.g.
 \begin{align}
@@
 \begin{align}
-  \diagram{triple_2_3} \equiv \diagram{triple_2_3_uvw}
-  + \diagram{triple_2_3_vwu} + \diagram{triple_2_3_wuv}.
+  \diagram{triple_2_3}
+  \equiv \f13 \sp{\diagram{triple_2_3_uvw}
+  + \diagram{triple_2_3_vwu} + \diagram{triple_2_3_wuv}}.
   \label{eq:triple_uvw_3}

### 2020-02-05 collective_sun/sun_perturbation.tex
simplified diagrams for triple product of three two-body operators

@@
   \P_0 \sp{\sum_{p,q} v_p w_q X_p Y_q} \P_0
-  \EQFS \diagram{single_body_2} X_1 Y_2
+  \EQFS \diagram{single_body_0} X_1 Y_2
   + \diagram{single_body_1} X_1 Y_1,
@@
 \begin{align}
-  \diagram{single_body_2}
-  = \diagram{single_body_2_o} - \diagram{single_body_2_x}
-  = \diagram{single_body_2_oo} - \diagram{single_body_1}
+  \diagram{single_body_0}
+  = \diagram{single_body_0_o} - \diagram{single_body_0_x}
+  = \diagram{single_body_0_oo} - \diagram{single_body_1}
   = - \v v \c\v w,
@@
     v_{k\ell} w_{pq} O_{k\ell} Q_{pq}} \P_0 \\
-  \EQFS \f14 \diagram{two_body_4} O_{1,2} Q_{3,4}
-  + \diagram{two_body_3} O_{1,2} Q_{1,3}
+  \EQFS \f14 \diagram{two_body_0} O_{1,2} Q_{3,4}
+  + \diagram{two_body_1} O_{1,2} Q_{1,3}
   + \f12 \diagram{two_body_2} O_{1,2} Q_{1,2},
@@
 \begin{align}
-  \diagram{two_body_3}
-  &= \diagram{two_body_3_o} - \diagram{two_body_3_x} \\
-  &= \diagram{two_body_3_oo} - \diagram{two_body_3_ox}
+  \diagram{two_body_1}
+  &= \diagram{two_body_1_o} - \diagram{two_body_1_x} \\
+  &= \diagram{two_body_1_oo} - \diagram{two_body_1_ox}
   - \diagram{two_body_2} \label{eq:mid_zero} \\
@@
 \begin{align}
-  \diagram{two_body_4}
-  &= \diagram{two_body_4_o} - \diagram{two_body_4_x} \\
-  &= \sp{\diagram{two_body_4_oo} - \diagram{two_body_4_ox}}
-  - 2 \sp{\diagram{two_body_3}} \\
-  &= \sp{-2 \diagram{two_body_3_o}}
-  - 2 \sp{\diagram{two_body_3_oo} - \diagram{two_body_2}} \\
-  &= -4 \diagram{two_body_3_oo} + 2 \diagram{two_body_2} \\
+  \diagram{two_body_0}
+  &= \diagram{two_body_0_o} - \diagram{two_body_0_x} \\
+  &= \sp{\diagram{two_body_0_oo} - \diagram{two_body_0_ox}}
+  - 2 \sp{\diagram{two_body_1}} \\
+  &= \sp{-2 \diagram{two_body_1_o}}
+  - 2 \sp{\diagram{two_body_1_oo} - \diagram{two_body_2}} \\
+  &= -4 \diagram{two_body_1_oo} + 2 \diagram{two_body_2} \\
   &= -4 \v v\c\v w + 4 \v{\m w}\c\v{\m w}.
@@
 \begin{align}
-  A_6 \equiv \f18 \diagram{triple_6},
-  &&
-  A_4 \equiv \diagram{triple_4_1} + \f32 \diagram{triple_4_0},
+  A_6 \equiv \f18 \diagram{triple_0},
+  &&
+  A_4 \equiv \diagram{triple_1} + \f32 \diagram{triple_01},
   \label{eq:triple_64}
@@
 \begin{align}
-  A_2 \equiv \f12 \diagram{triple_2_0} + 3 \diagram{triple_2_1}
-  + \f34 \diagram{triple_2_2} + 3 \diagram{triple_2_3},
+  A_2 \equiv \f12 \diagram{triple_2} + 3 \diagram{triple_11}
+  + \f34 \diagram{triple_02} + 3 \diagram{triple_011},
   \label{eq:triple_2}
@@
 \begin{align}
-  \diagram{triple_6} \equiv \diagram{triple_6_uvw},
+  \diagram{triple_0} \equiv \diagram{triple_0_uvw},
   \label{eq:triple_uvw}
@@
 \begin{align}
-  \diagram{triple_2_3}
-  \equiv \f13 \sp{\diagram{triple_2_3_uvw}
-  + \diagram{triple_2_3_vwu} + \diagram{triple_2_3_wuv}}.
+  \diagram{triple_011}
+  \equiv \f13 \sp{\diagram{triple_011_uvw}
+  + \diagram{triple_011_vwu} + \diagram{triple_011_wuv}}.
   \label{eq:triple_uvw_3}
@@
 of labels, as in \eqref{eq:triple_uvw}, while all other unlabeled
-diagrams have three assignments, as in \eqref{eq:triple_uvw_3}.  We
-can write \eqref{eq:triple_multi} in terms of collective operators as
+diagrams have three assignments, as in \eqref{eq:triple_uvw_3}.
+
+Having identified diagrammatic representations of the coefficients
+$A_6,A_4,A_2$, we now need to simplify all diagrams to reduce their
+computational complexity, as discussed in Section \ref{sec:diagrams}.
+Eliminating all filled dots in regions that address only one tensor
+yields:
+\begin{align}
+  A_6 = D_1 - D_2 + D_3 - D_4,
+\end{align}
+\begin{align}
+  A_4 = D_2 - 2 D_3 + 3 D_4,
+  &&
+  A_2 = D_3 - 3 D_4
+\end{align}
+where
+\begin{align}
+  D_1 \equiv \f18 \diagram{triple_0_o6},
+  &&
+  D_2 \equiv \f32 \diagram{triple_01_o4}
+  - 2 \diagram{triple_1_o3},
+  &&
+  D_4 \equiv \diagram{triple_0111}.
+\end{align}
+\begin{align}
+  D_3 \equiv 3 \diagram{triple_011_o2}
+  + \f34 \diagram{triple_02_o2}
+  - 6 \diagram{triple_11_o}
+  + 2 \diagram{triple_2}.
+\end{align}
+We can also write the product $\X$ in \eqref{eq:triple_multi} using
+collective operators as
 \begin{align}

### 2020-02-05 collective_sun/scripts/make_diagrams.py

@@ -207,17 +207,17 @@ add_triple_labels(["w_1","w_2","w_3"], pads = [ 0.25, 0.5, 0.55 ])
 plt.tight_layout(pad = 0)
 plt.savefig(fig_dir + "example_triple.pdf")
 
-make_triple_diagram(dots, {"a" : "o", "ac" : "o"}, figsize = (1,1))
+make_triple_diagram(dots, { "a" : "o", "ac" : "o" }, figsize = (1,1))
 add_triple_labels(["w_1","w_2","w_3"], pads = [ 0.25, 0.5, 0.55 ])
 plt.tight_layout(pad = 0)
 plt.savefig(fig_dir + "example_triple_oo_2.pdf")
 
-make_triple_diagram(dots, {"ac" : "o", "bc" : "o"}, figsize = (1,1))
+make_triple_diagram(dots, { "ac" : "o", "bc" : "o" }, figsize = (1,1))
 add_triple_labels(["w_1","w_2","w_3"], pads = [ 0.25, 0.5, 0.55 ])
 plt.tight_layout(pad = 0)
 plt.savefig(fig_dir + "example_triple_oo_1.pdf")
 
-make_triple_diagram(dots, {"abc" : "x", "ac" : "x"}, figsize = (1,1))
+make_triple_diagram(dots, { "abc" : "x", "ac" : "x" }, figsize = (1,1))
 add_triple_labels(["w_1","w_2","w_3"], pads = [ 0.25, 0.5, 0.55 ])
 plt.tight_layout(pad = 0)
 plt.savefig(fig_dir + "example_triple_xx_1.pdf")
@@ -227,28 +227,28 @@ plt.close("all")
 ### example diagrams for open dot elimination
 
 dots = { "a" : 1, "abc" : 1, "ac" : 1, "bc" : 2 }
-make_triple_diagram(dots, {"a" : "o"}, figsize = (1,1))
+make_triple_diagram(dots, { "a" : "o" }, figsize = (1,1))
 add_triple_labels(["w_1","w_2","w_3"], pads = [ 0.25, 0.5, 0.55 ])
 plt.tight_layout(pad = 0)
 plt.savefig(fig_dir + "example_triple_o_top.pdf")
 
-make_triple_diagram(dots, {"a" : "x"}, figsize = (1,1))
+make_triple_diagram(dots, { "a" : "x" }, figsize = (1,1))
 add_triple_labels(["w_1","w_2","w_3"], pads = [ 0.25, 0.5, 0.55 ])
[…]

### 2020-02-05 collective_sun/sun_perturbation.tex
explicitly provide diagram factors in the triple product in terms of the two-body coupling matrices

@@
 not sum over the values of indices in any tensor that contains that
-index; that is (hint: pay attention to the second sum on each line),
+index; that is (pay attention to the second sum in each line),
 \begin{align}
@@
 \end{align}
-We can also write the product $\X$ in \eqref{eq:triple_multi} using
-collective operators as
+Defining, for each of $m\in\set{u,v,w}$,
+\begin{align}
+  \m m \equiv \sum_{p,q} m_{pq} \op{p}{q},
+  &&
+  \v{\m m} \equiv \sum_{p<q} m_{pq} \ket{\set{p,q}},
+  &&
+  \v m \equiv \sum_p m_p \ket{p},
+  &&
+  m_p \equiv \sum_q m_{pq},
+  &&
+  \col{m} \equiv \sum_{p<q} m_{pq},
+\end{align}
+we can expand
+\begin{align}
+  D_1 = \col{u}\,\col{v}\,\col{w},
+  &&
+  D_2 = \col{u}\,\v v\c\v w + \col{v}\,\v w\c\v u
+  + \col{w}\,\v u\c\v v - 2 \sum_p u_p v_p w_p,
+  &&
+  D_4 = \tr\p{\m u \c \m v \c \m w},
+\end{align}
+\begin{multline}
+  D_3 = \v u \c\m v\c\v w + \v v \c\m w\c\v u + \v w \c\m u\c\v v
+  + \col{u}\,\v{\m v}\c\v{\m w} + \col{v}\,\v{\m w}\c\v{\m u}
+  + \col{w}\,\v{\m u}\c\v{\m v} \\
+  - 2 \sum_{p,q} \p{u_p v_{pq} w_{pq}
+    + v_p w_{pq} u_{pq} + w_p u_{pq} v_{pq}}
+  + 4 \sum_{p<q} u_{pq} v_{pq} w_{pq}.
+\end{multline}
+Finally, we also write the product $\X$ in \eqref{eq:triple_multi}
+using the collective operator $\col{X} \equiv \sum_p X_p$:
 \begin{align}

### 2020-02-05 collective_sun/sun_perturbation.tex
minor wording changes

@@
 \end{align}
-where a diagram with unlabeled circles denotes an average over all
-label assignments that result in a distinct diagram, e.g.
+where we have neglected scalar terms without physical consequence, and
+a diagram with unlabeled circles denotes an average over all distinct
+label assignments, e.g.
 \begin{align}
@@
 \end{multline}
-Finally, we also write the product $\X$ in \eqref{eq:triple_multi}
-using the collective operator $\col{X} \equiv \sum_p X_p$:
+Finally, we can also write the product $\X$ in \eqref{eq:triple_multi}
+in terms of the collective operator $\col{X} \equiv \sum_p X_p$:
 \begin{align}

### 2020-02-05 collective_sun/sun_perturbation.tex
renamed some variables

@@
 \begin{align}
-  A_6 = D_1 - D_2 + D_3 - D_4,
-\end{align}
-\begin{align}
-  A_4 = D_2 - 2 D_3 + 3 D_4,
-  &&
-  A_2 = D_3 - 3 D_4
+  A_6 = D_0 - D_1 + D_2 - D_3,
+  &&
+  A_4 = D_1 - 2 D_2 + 3 D_3,
+  &&
+  A_2 = D_2 - 3 D_3
 \end{align}
@@
 \begin{align}
-  D_1 \equiv \f18 \diagram{triple_0_o6},
-  &&
-  D_2 \equiv \f32 \diagram{triple_01_o4}
+  D_0 \equiv \f18 \diagram{triple_0_o6},
+  &&
+  D_1 \equiv \f32 \diagram{triple_01_o4}
   - 2 \diagram{triple_1_o3},
   &&
-  D_4 \equiv \diagram{triple_0111}.
-\end{align}
-\begin{align}
-  D_3 \equiv 3 \diagram{triple_011_o2}
+  D_3 \equiv \diagram{triple_0111}.
+\end{align}
+\begin{align}
+  D_2 \equiv 3 \diagram{triple_011_o2}
   + \f34 \diagram{triple_02_o2}
@@
 \begin{align}
-  D_1 = \col{u}\,\col{v}\,\col{w},
-  &&
-  D_2 = \col{u}\,\v v\c\v w + \col{v}\,\v w\c\v u
+  D_0 = \col{u}\,\col{v}\,\col{w},
+  &&
+  D_1 = \col{u}\,\v v\c\v w + \col{v}\,\v w\c\v u
   + \col{w}\,\v u\c\v v - 2 \sum_p u_p v_p w_p,
   &&
-  D_4 = \tr\p{\m u \c \m v \c \m w},
+  D_3 = \tr\p{\m u \c \m v \c \m w},
 \end{align}
 \begin{multline}
-  D_3 = \v u \c\m v\c\v w + \v v \c\m w\c\v u + \v w \c\m u\c\v v
+  D_2 = \v u \c\m v\c\v w + \v v \c\m w\c\v u + \v w \c\m u\c\v v
   + \col{u}\,\v{\m v}\c\v{\m w} + \col{v}\,\v{\m w}\c\v{\m u}

### 2020-02-05 collective_sun/sun_perturbation.tex
changed definition of unlabeled diagram to be a sum (rather than an average) over distinct label assignments

@@
   &&
-  A_4 \equiv \diagram{triple_1} + \f32 \diagram{triple_01},
+  A_4 \equiv \diagram{triple_1} + \f12 \diagram{triple_01},
   \label{eq:triple_64}
@@
 \begin{align}
-  A_2 \equiv \f12 \diagram{triple_2} + 3 \diagram{triple_11}
-  + \f34 \diagram{triple_02} + 3 \diagram{triple_011},
+  A_2 \equiv \f12 \diagram{triple_2} + \diagram{triple_11}
+  + \f14 \diagram{triple_02} + \diagram{triple_011},
   \label{eq:triple_2}
@@
 where we have neglected scalar terms without physical consequence, and
-a diagram with unlabeled circles denotes an average over all distinct
-label assignments, e.g.
-\begin{align}
-  \diagram{triple_0} \equiv \diagram{triple_0_uvw},
+a diagram with unlabeled circles denotes a sum over all distinct label
+assignments, e.g.
+\begin{align}
+  \diagram{triple_1} \equiv \diagram{triple_1_uvw},
   \label{eq:triple_uvw}
@@
   \diagram{triple_011}
-  \equiv \f13 \sp{\diagram{triple_011_uvw}
-  + \diagram{triple_011_vwu} + \diagram{triple_011_wuv}}.
+  \equiv \diagram{triple_011_uvw}
+  + \diagram{triple_011_vwu} + \diagram{triple_011_wuv}.
   \label{eq:triple_uvw_3}
@@
   &&
-  A_2 = D_2 - 3 D_3
+  A_2 = D_2 - 3 D_3,
 \end{align}
@@
   &&
-  D_1 \equiv \f32 \diagram{triple_01_o4}
-  - 2 \diagram{triple_1_o3},
-  &&
-  D_3 \equiv \diagram{triple_0111}.
-\end{align}
-\begin{align}
-  D_2 \equiv 3 \diagram{triple_011_o2}
-  + \f34 \diagram{triple_02_o2}
-  - 6 \diagram{triple_11_o}
+  D_1 \equiv \f12 \diagram{triple_01_o4} - 2 \diagram{triple_1_o3},
+  &&
+  D_3 \equiv \diagram{triple_0111},
+\end{align}
+\begin{align}
+  D_2 \equiv \diagram{triple_011_o2}
+  + \f14 \diagram{triple_02_o2}
+  - 2 \diagram{triple_11_o}
   + 2 \diagram{triple_2}.

### 2020-02-05 collective_sun/scripts/make_diagrams.py

@@ -368,16 +368,16 @@ plt.close("all")
 make_triple_diagram({ "a" : 2, "b" : 2, "c" : 2 })
 plt.savefig(fig_dir + "triple_0.pdf")
 
-make_triple_diagram({ "a" : 2, "b" : 2, "c" : 2 }, figsize = (0.9,0.7))
-add_triple_labels(["u","v","w"])
-plt.tight_layout(pad = 0.05)
-plt.savefig(fig_dir + "triple_0_uvw.pdf")
-
 plt.close("all")
 
 # 4-point diagrams
 make_triple_diagram({ "a" : 1, "b" : 1, "c" : 1, "abc" : 1 })
 plt.savefig(fig_dir + "triple_1.pdf")
+
+make_triple_diagram({ "a" : 1, "b" : 1, "c" : 1, "abc" : 1 }, figsize = (0.9,0.7))
+add_triple_labels(["u","v","w"])
+plt.tight_layout(pad = 0.05)
+plt.savefig(fig_dir + "triple_1_uvw.pdf")
 
 make_triple_diagram({ "a" : 2, "b" : 1, "c" : 1, "bc" : 1 })
 plt.savefig(fig_dir + "triple_01.pdf")

### 2020-02-05 collective_sun/sun_perturbation.tex
fully simplify diagrams in the triple product

@@
   &= -4 \diagram{two_body_1_oo} + 2 \diagram{two_body_2} \\
-  &= -4 \v v\c\v w + 4 \v{\m w}\c\v{\m w}.
+  &= -4 \v v\c\v w + 4 \v{\m v}\c\v{\m w}.
 \end{align}
@@
 computational complexity, as discussed in Section \ref{sec:diagrams}.
-Eliminating all filled dots in regions that address only one tensor
-yields:
+Simplifying all these diagrams by hand is tedious, but the process of
+eliminating filled dots and crosses from a diagram is algorithmic
+enough to execute on a computer, doing which we find
 \begin{align}
@@
 \begin{align}
-  D_0 \equiv \f18 \diagram{triple_0_o6},
-  &&
-  D_1 \equiv \f12 \diagram{triple_01_o4} - 2 \diagram{triple_1_o3},
-  &&
-  D_3 \equiv \diagram{triple_0111},
-\end{align}
-\begin{align}
-  D_2 \equiv \diagram{triple_011_o2}
-  + \f14 \diagram{triple_02_o2}
+  D_0 \equiv \f18 \diagram{triple_0_o},
+  &&
+  D_1 \equiv \f12 \diagram{triple_01_o} - 2 \diagram{triple_1_o},
+  &&
+  D_3 \equiv \diagram{triple_0111_o},
+\end{align}
+\begin{align}
+  D_2 \equiv \diagram{triple_011_o}
+  + \f14 \diagram{triple_02_o}
   - 2 \diagram{triple_11_o}
-  + 2 \diagram{triple_2}.
+  + 2 \diagram{triple_2_o}.
 \end{align}

### 2020-02-05 collective_sun/scripts/make_diagrams.py

@@ -71,14 +71,17 @@ region_mids["ac"] = rotate("b") @ region_mids["bc"]
 # define constructor for diagram
 
 def place_dots(region, num, markers = None, dot_sep = None,
-               region_mids = region_mids, markersize = 3):
+               region_mids = region_mids, markersize = 3,
+               default_marker = None):
+    if default_marker is None:
+        default_marker = "."
     if dot_sep == None:
         dot_sep = 1
     dot_sep *= radius / 3
     if markers is None:
-        markers = "." * num
+        markers = default_marker * num
     if len(markers) < num:
-        markers += "." * (num-len(markers))
+        markers += default_marker * (num-len(markers))
 
     vecs = []
 
@@ -116,7 +119,8 @@ def place_dots(region, num, markers = None, dot_sep = None,
         plt.plot(*pos, "k"+marker, markersize = size, fillstyle = fill_style)
 
 def make_triple_diagram(region_dots, markers = {}, figsize = figsize,
-                        dot_sep = None, pad = radius/20):
+                        dot_sep = None, pad = radius/20,
+                        default_marker = None):
     fig = plt.figure(figsize = figsize)
     plt.axes().set_aspect("equal")
 
@@ -146,7 +150,8 @@ def make_triple_diagram(region_dots, markers = {}, figsize = figsize,
 
     # plot dots
     for region, num in region_dots.items():
-        place_dots(region, num, markers.get(region), dot_sep)
+        place_dots(region, num, markers.get(region), dot_sep,
+                   default_marker = default_marker)
 
     # trim figure and return handle
     plt.axis("off")
@@ -154,7 +159,8 @@ def make_triple_diagram(region_dots, markers = {}, figsize = figsize,
     return fig
 
 def make_double_diagram(region_dots, markers = {}, size_x = figsize[1],
-                        dot_sep = None, pad = radius/20):
+                        dot_sep = None, pad = radius/20,
[…]

### 2020-02-05 collective_sun/scripts/simplify_product_diagrams.py

@@ -92,26 +92,25 @@ def permute_diagram(diagram, permutation):
     return { tuple(sorted( permutation[idx] for idx in region )) : markers
              for region, markers in diagram.items() }
 
-# reduce a diagram by eliminating filled dots and crosses in single-tensor regions
+# reduce a diagram by eliminating filled dots and crosses
 def reduce_diagram(diagram):
     diagram = { region : markers if type(markers) is tuple else (markers,0,0)
                 for region, markers in diagram.items()
                 if markers != 0 and markers != (0,0,0) }
 
-    def _net_lone_markers(mm):
-        return sum([ markers[mm]
-                     for region, markers in diagram.items()
-                     if len(region) == 1 ])
+    def _net_markers(mm):
+        return sum([ markers[mm] for region, markers in diagram.items() ])
 
-    # first, cover the simpler case of a diagram without any crosses
-    if _net_lone_markers(2) == 0:
-        if _net_lone_markers(0) == 0:
-            cross_free_diagram = { region : markers[:2]
-                                   for region, markers in diagram.items() }
-            return diagram_vec(cross_free_diagram)
+    # first, cover the simpler case of a diagram without any crosses,
+    #        in which we eliminate any filled dots that we find
+    if _net_markers(2) == 0:
+        # if there are no crosses and no filled dots, then we are done simplifying
+        if _net_markers(0) == 0:
+            empty_dot_diagram = { region : markers[1]
+                                  for region, markers in diagram.items() }
+            return diagram_vec(empty_dot_diagram)
 
         for region, markers in diagram.items():
-            if len(region) != 1: continue
             if markers[0] == 0: continue
             empty_copy = copy.deepcopy(diagram)
             cross_copy = copy.deepcopy(diagram)
@@ -122,19 +121,18 @@ def reduce_diagram(diagram):
     # otherwise, eliminate any crosses in a diagram
     else:
         for region, markers in diagram.items():
-            if len(region) != 1: continue
             if markers[2] == 0: continue
 
-            this_primary_set = region[0]
             def _take_from_region(other_region):
                 other_markers = diagram[other_region]
                 mutiplicity = other_markers[0]
+                if mutiplicity == 0: return
 
                 new_diagram = copy.deepcopy(diagram)
-                new_diagram[(this_primary_set,)] = markers[:2] + ( markers[2]-1, )
[…]

### 2020-02-06 collective_sun/sun_perturbation.tex
final changes to diagram definitions

@@
 \begin{align}
-  \diagram{example_triple},
+  \diagram{example_123},
   \label{eq:venn_diagram}
@@
 The simplest diagrams have $g_S$ dots (``$\bullet$'') at the
-intersection of the circles $j\in S$, and are proportional to the
+intersection of the circles $j\in S$, and are simply equal to the
 associated scalar $w\p{g}$ defined in \eqref{eq:diagram_factors}, e.g.
 \begin{align}
-  \diagram{example_triple}
-  \equiv \sp{\prod_{S\in\PP\p{\ZZ_p}} g_S!} w\p{g}
-  = \sum_{\substack{a,b,c,d,e\in\ZZ_N\\
+  \diagram{example_123}
+  \equiv \sum_{\substack{a,b,c,\in\ZZ_N\\\p{d,e}\in\C_N\p{2}\\
       \abs{\set{a,b,c,d,e}}=5}}
   w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}.
-  \label{eq:diagram_definition}
-\end{align}
-The proportionality factor accounts for the number of ways to permute
-the indices within each region $S\in\PP\p{\ZZ_p}$.  Including this
-symmetry factor in the definition of a diagram is a matter of choice
-that will ultimately make it easier to work with diagrams.  A diagram
-such at that in \eqref{eq:diagram_definition} with $\abs{g}$ dots
-nominally takes $\O\p{N^{\abs{g}}}$ time to compute.  In practice,
-this computational cost can be prohibitive for performing numerical
-simulations of systems of any appreciable size.  If possible, we
-therefore wish to simplify diagrams in order to reduce their
-computational complexity.  To this end, we first need to define simple
-modifications of diagrams in which dots may be replaced by empty dots
+\end{align}
+Such a diagram with $\abs{g}$ dots nominally takes $\O\p{N^{\abs{g}}}$
+time to compute.  In practice, this computational cost can be
+prohibitive for performing numerical simulations of systems with any
+appreciable size.  We therefore wish to simplify diagrams in order to
+reduce their computational cost as much as possible.  Our general
+strategy will be to replace constrained sums that make all indices
+interdependent (such as $\abs{\set{a,b,c,d,e}}=5$ above) by
+unconstrained sums that can be carried out in sequence.  In order to
+represent different constraints, we need to define diagrams in which
+filled dots (``$\bullet$'') may be replaced by empty dots
 (``$\circ$'') or crosses (``$\bm\times$'').
 
-In any diagram (with filled dots, empty dots, or crosses), the values
-of indices associated with filled dots are always summed as if the
-diagram had only dots, without regard for all other indices.  Empty
-dots in a diagram then correspond to indices that are summed over all
-of $\ZZ_N$ independently and without additional constraints, e.g.
-\begin{align}
-  \diagram{example_triple_oo_2} \equiv
-  \sum_{a,c\in\ZZ_N}
-  \sum_{\substack{b,d,e\in\ZZ_N\\\abs{\set{b,d,e}}=3}}
-  w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}.
-\end{align}
-Crosses, meanwhile, correspond to indices that are {\it maximally
-  dependent} on filled dots: ``cross indices'' are summed {\it only}
-over the values of ``filled dot indices'', e.g.
-\begin{align}
-  \diagram{example_triple_xx_1}
-  \equiv \sum_{\substack{a,d,e\in\ZZ_N\\\abs{\set{a,d,e}}=3}}
-  \sum_{b,c\in\set{a,d,e}}
-  w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}.
-\end{align}
-Empty dots and crosses allow us to decompose diagrams into sums of
-different diagrams that (may) require less time to evaluate.  We will
-essentially use two tricks to decompose diagrams: eliminating filled
-dots, and eliminating crosses.
+If a region of a diagram contains $F$ filled dots, those dots
+correspond to indices that sum over $\C_N\p{F}$ with the constraint
+that they are not equal to any other filled dots in a diagram.  $E$
+empty dots in a region, meanwhile, correspond to indices that are
+summed over $\C_N\p{E}$ without any additional constraints, e.g.
+\begin{align}
+  \diagram{example_o}
+  \equiv \sum_{\substack{\p{a,b}\in\C_N\p{2}\\e\in\ZZ_N\\
+      \abs{\set{a,b,e}}=3}}
+  \sum_{\substack{\p{c,d}\in\C_N\p{2}\\f\in\ZZ_N}}
+  v\p{a,b,c,d,e} w\p{e,f}.
+\end{align}
+Empty dots are therefore entirely independent of filled dots.
+Crosses, meanwhile, are {\it entirely dependent} on filled dots:
+``cross indices'' are summed {\it only} over the values of ``filled
+dot indices'', e.g.
+\begin{align}
+  \diagram{example_x}
+  \equiv \sum_{\substack{a,e\in\ZZ_N\\\abs{\set{a,e}}=2}}
+  \sum_{b,c\in\set{a,e}} \sum_{d\in\ZZ_N}
+  u\p{a,b,c} v\p{b,d,e} w\p{b,c,e}.
+\end{align}
+Empty dots and crosses allow us to decompose diagrams with a high
+computational cost into different diagrams with a lower computational
+cost.  We will essentially use two tricks to decompose diagrams:
+eliminating filled dots, and eliminating crosses.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Decomposing diagrams}
 
@@
 \begin{align}
-  \diagram{example_triple}
-  = \diagram{example_triple_o_top} - \diagram{example_triple_x_top},
-  \label{eq:example_decomposition}
+  \diagram{example_elim}
+  = \diagram{example_elim_o}
+  - \diagram{example_elim_x}
+  \label{eq:example_elim}
 \end{align}
@@
 \begin{align}
-  \diagram{example_triple}
-  \equiv \sum_{\substack{a,b,c,d,e\in\ZZ_N\\\abs{\set{a,b,c,d,e}}=5}}
-  w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e},
-\end{align}
-\begin{align}
-  \diagram{example_triple_x_top}
-  \equiv \sum_{\substack{b,c,d,e\in\ZZ_N\\\abs{\set{b,c,d,e}}=4}}
-  \sum_{a\in\set{b,c,d,e}}
-  w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e},
-\end{align}
-\begin{align}
-  \diagram{example_triple_o_top}
-  \equiv \sum_{\substack{b,c,d,e\in\ZZ_N\\\abs{\set{b,c,d,e}}=4}}
-  w_1\p{\circ,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e},
-  \label{eq:example_triple_x_top}
-\end{align}
-with
-\begin{align}
-  w_1\p{\circ,b,c} \equiv \sum_{a\in\ZZ_N} w_1\p{a,b,c},
-\end{align}
-The decomposition in \eqref{eq:example_decomposition} essentially
-breaks up a sum over $a\in\ZZ_N\setminus\set{b,c,d,e}$ into a
-difference of sums over $a\in\ZZ_N$ and $a\in\set{b,c,d,e}$.  The sum
-over $a\in\ZZ_N$ to compute $w_1\p{\circ,b,c}$ has $\O\p{N^3}$ cost,
-and can be performed prior to computing the diagram in
-\eqref{eq:example_triple_x_top}.  The decomposition in
-\eqref{eq:example_decomposition} thereby reduces the calculation of
-one $\O\p{N^5}$ diagram to the calculation of two $\O\p{N^4}$
-diagrams.  A similar reduction by a factor of $\sim N$ can be achieved
-by decomposing any dot in a region $S\in\PP\p{\ZZ_p}$ addressing only
-one tensor, i.e.~with $\abs{S}=1$.
-
-While it is possible, in principle, to derive rules for decomposing
-filled dots in a diagram that contains crosses, it is much easier to
-simply eliminate crosses from a diagram altogether.  To eliminate a
-cross from a diagram, we first note that the corresponding index need
-not sum over the values of indices in any tensor that contains that
-index; that is (pay attention to the second sum in each line),
-\begin{align}
-  \diagram{example_triple_x_top_elim}
-  &\equiv \sum_{\substack{b,c,d,e,f\in\ZZ_N\\\abs{\set{b,c,d,e,f}}=5}}
-  \sum_{a\in\set{b,c,d,e,f}}
-  w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,f} \\
-  &= \sum_{\substack{b,c,d,e,f\in\ZZ_N\\\abs{\set{b,c,d,e,f}}=5}}
-  \sum_{a\in\set{d,e,f}}
-  w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,f}.
-  \label{eq:x_elim}
+  \diagram{example_elim}
+  \equiv \sum_{\substack{a,b,c\in\ZZ_N\\\abs{\set{a,b,c}}=2}}
+  v\p{a,b} w\p{b,c},
+\end{align}
+\begin{align}
+  \diagram{example_elim_o}
+  \equiv \sum_{\substack{b,c\in\ZZ_N\\\abs{\set{b,c}}=2}}
+  v\p{\circ,b} w\p{b,c},
+  &&
+  v\p{\circ,b} \equiv \sum_{a\in\ZZ_N} v\p{a,b},
+  \label{eq:example_elim_o}
+\end{align}
+\begin{align}
+  \diagram{example_elim_x}
+  \equiv \sum_{\substack{b,c\in\ZZ_N\\\abs{\set{b,c}}=2}}
+  \sum_{a\in\set{b,c}} v\p{a,b} w\p{b,c}.
+\end{align}
+The decomposition in \eqref{eq:example_elim} essentially breaks
+up a sum over $a\in\ZZ_N\setminus\set{b,c}$ into a difference of sums
+over $a\in\ZZ_N$ and $a\in\set{b,c}$.  The sum over $a\in\ZZ_N$ to
+compute $v\p{\circ,b}$ has $\O\p{N^2}$ cost, and can be performed
+prior to evaluating the diagram in \eqref{eq:example_elim_o}.
+The decomposition in \eqref{eq:example_elim} therefore splits
+one $\O\p{N^3}$ into two $\O\p{N^2}$ diagrams.
+
+After splitting a filled dot into an empty dot and a cross, the next
+step is to immediately eliminate the cross.  To eliminate a cross from
+a diagram, we first note that the corresponding index can ignore other
+indices on tensors that contain the index; that is (pay attention to
+the sum over $a$),
+\begin{align}
+  \diagram{example_elim_x}
+  \equiv \sum_{\substack{b,c\in\ZZ_N\\\abs{\set{b,c}}=2}}
+  \sum_{a\in\set{b,c}} v\p{a,b} w\p{b,c}
+  = \sum_{\substack{b,c\in\ZZ_N\\\abs{\set{b,c}}=2}}
+  \sum_{a\in\set{c}} v\p{a,b} w\p{b,c}.
 \end{align}
 For each remaining index addressed by a cross, meanwhile, the cross
-simply forces the corresponding indices to be equal, so
-\begin{align}
-  \diagram{example_triple_x_top_elim}
-  = 2 \diagram{example_triple_x_lft_elim}
-  + \diagram{example_triple_x_rht_elim},
-\end{align}
-where the factor of 2 appears simply to account for both of
-$a\in\set{d,e}$ in \eqref{eq:x_elim}.  The rules derived here for
-replacing filled dots and crosses will be used in the following
-sections to simplify diagrams and minimize the computational cost of
-their evaluation.
+simply forces the corresponding indices to be equal, which is
+equivalent to moving a filled dot into a different region:
+\begin{align}
+  \diagram{example_elim_x}
+  = 2 \diagram{example_elim_x_full}
+  = 2 \sum_{b,c\in\C_N\p{2}} v\p{c,b} w\p{b,c}.
+  \label{eq:example_elim_final}
+\end{align}
+The factor of 2 accounts for the fact that the left diagram sums over
+$b,c\in\ZZ_N$, while the right diagram sums over
+$\p{b,c}\in\C_N\p{2}$.  We can generally repeat the above process of
+eliminating filled dots and crosses until only empty dots remain in a
+diagram.  The only complication in carrying out this process is that
+we have to manually keep track of symmetry factors at each step, such
+as the factor of 2 in \eqref{eq:example_elim_final}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Symmetry factors}
+
+Here we discuss the rules governing symmetry factors that appear when
+decomposing diagrams.  In general, eliminating a filled dot from a
+region with $F$ filled dots makes a diagram pick up a factor of $1/F$,
+e.g.
+\begin{align}
+  \diagram{example_sym}
+  = \f14 \diagram{example_sym_o}
+  - \f14 \diagram{example_sym_x}.
+\end{align}
+This factor can be derived by replacing the sum over $\C_N\p{F}$ for
+the filled dots within the region by a sum over $\ZZ_N^{F}$, which
+picks up a factor of $1/F!$ to account for the number of ways to
+permute $F$ values.  Setting aside one filled dot for replacement by
+an empty dot and a cross, the sum over $\ZZ_N^{F-1}$ for the remaining
+dots can be changed back to a sum over $\C_N\p{F-1}$, picking up a
+factor of $\p{F-1}!$.  The overall factor picked up throughout this
+procedure is therefore $\p{F-1}!/F!=1/F$.  Similarly, adding an empty
+dot to a region with $E$ empty dots makes the diagram pick up a factor
+of $\p{E+1}!/E!=E+1$, so
+\begin{align}
+  \diagram{example_sym_o}
+  = \f23 \diagram{example_sym_oo}
+  - \f13 \diagram{example_sym_ox}.
+\end{align}
+Finally, eliminating a cross by moving a filled dot from an ``old''
+region with $F_{\t{old}}$ filled dots into a ``new'' region with
+$F_{\t{new}}$ filled dots picks up an overall factor of $F_{\t{new}}$,
+so
+\begin{align}
+  \diagram{example_sym_x}
+  = 2 \diagram{example_sym_x_elim},
+\end{align}
+where the factor of $1/F_{\t{old}}$ that would be acquired from
+removing a dot from the old region is canceled out by a factor of
+$F_{\t{old}}$ that accounts for the number of ways to move a dot from
+the old region to the new one.
 
@@
     v_{k\ell} w_{pq} O_{k\ell} Q_{pq}} \P_0 \\
-  \EQFS \f14 \diagram{two_body_0} O_{1,2} Q_{3,4}
+  \EQFS \diagram{two_body_0} O_{1,2} Q_{3,4}
   + \diagram{two_body_1} O_{1,2} Q_{1,3}
-  + \f12 \diagram{two_body_2} O_{1,2} Q_{1,2},
+  + \diagram{two_body_2} O_{1,2} Q_{1,2},
   \label{eq:POQP_start}
@@
 \begin{align}
-  \diagram{two_body_2} = \sum_{p,q\in\ZZ_N} v_{pq} w_{pq}
-  = 2 \sum_{p,q\in\C_N\p{2}} v_{pq} w_{pq}
+  \diagram{two_body_2} \equiv \sum_{p,q\in\C_N\p{2}} v_{pq} w_{pq}
   = \v{\m v} \c\v{\m w},
@@
   \diagram{two_body_1}
-  &= \diagram{two_body_1_o} - \diagram{two_body_1_x} \\
-  &= \diagram{two_body_1_oo} - \diagram{two_body_1_ox}
-  - \diagram{two_body_2} \label{eq:mid_zero} \\
+  &= \sp{\diagram{two_body_1_o}} - \sp{\diagram{two_body_1_x}} \\
+  &= \sp{\diagram{two_body_1_oo} - \diagram{two_body_1_ox}}
+  - \sp{2\diagram{two_body_2}} \label{eq:mid_zero} \\
   &= \v v \c \v w - 2 \v{\m v} \c \v{\m w},
@@
   \diagram{two_body_0}
-  &= \diagram{two_body_0_o} - \diagram{two_body_0_x} \\
-  &= \sp{\diagram{two_body_0_oo} - \diagram{two_body_0_ox}}
-  - 2 \sp{\diagram{two_body_1}} \\
-  &= \sp{-2 \diagram{two_body_1_o}}
-  - 2 \sp{\diagram{two_body_1_oo} - \diagram{two_body_2}} \\
-  &= -4 \diagram{two_body_1_oo} + 2 \diagram{two_body_2} \\
-  &= -4 \v v\c\v w + 4 \v{\m v}\c\v{\m w}.
+  &= \f12 \sp{\diagram{two_body_0_o}}
+  - \f12 \sp{\diagram{two_body_0_x}} \\
+  &= \f12 \sp{\diagram{two_body_0_oo} - \diagram{two_body_0_ox}}
+  - \f12 \sp{\diagram{two_body_1}} \\
+  &= \f12 \sp{-\diagram{two_body_1_o}}
+  - \f12 \p{\v v\c\v w - 2\v{\m v}\c\v{\m w}} \\
+  &= -\v v\c\v w + \v{\m v}\c\v{\m w}.
 \end{align}
@@
 \begin{align}
-  A_6 \equiv \f18 \diagram{triple_0},
-  &&
-  A_4 \equiv \diagram{triple_1} + \f12 \diagram{triple_01},
+  A_6 \equiv \diagram{triple_0},
+  &&
+  A_4 \equiv \diagram{triple_01} + \diagram{triple_1},
   \label{eq:triple_64}
@@
 \begin{align}
-  A_2 \equiv \f12 \diagram{triple_2} + \diagram{triple_11}
-  + \f14 \diagram{triple_02} + \diagram{triple_011},
+  A_2 \equiv \diagram{triple_011} + \diagram{triple_02}
+  + \diagram{triple_11} + \diagram{triple_2},
   \label{eq:triple_2}
@@
 \begin{align}
-  D_0 \equiv \f18 \diagram{triple_0_o},
-  &&
-  D_1 \equiv \f12 \diagram{triple_01_o} - 2 \diagram{triple_1_o},
+  D_0 \equiv \diagram{triple_0_o},
+  &&
+  D_1 \equiv \diagram{triple_01_o} - 2 \diagram{triple_1_o},
   &&
@@
   D_2 \equiv \diagram{triple_011_o}
-  + \f14 \diagram{triple_02_o}
+  + \diagram{triple_02_o}
   - 2 \diagram{triple_11_o}
-  + 2 \diagram{triple_2_o}.
+  + 4 \diagram{triple_2_o}.
 \end{align}

### 2020-02-06 collective_sun/scripts/make_diagrams.py

@@ -96,7 +96,13 @@ def place_dots(region, num, markers = None, dot_sep = None,
         sep_vec = np.array([0,dot_sep])
         vecs = [ rotate(jj) @ sep_vec / np.sqrt(3) for jj in range(3) ]
 
-    elif num > 3:
+    elif num == 4:
+        vecs = [ [ -dot_sep/2, +dot_sep/2 ],
+                 [ +dot_sep/2, +dot_sep/2 ],
+                 [ +dot_sep/2, -dot_sep/2 ],
+                 [ -dot_sep/2, -dot_sep/2 ] ]
+
+    elif num > 4:
         print("capability to place more than two dots not implemented")
         exit()
 
@@ -115,7 +121,7 @@ def place_dots(region, num, markers = None, dot_sep = None,
         if marker == "+":
             size += 1
 
-        pos = point + vec
+        pos = np.array(point) + np.array(vec)
         plt.plot(*pos, "k"+marker, markersize = size, fillstyle = fill_style)
 
 def make_triple_diagram(region_dots, markers = {}, figsize = figsize,
@@ -210,77 +216,97 @@ def add_double_labels(labels, pads = [ 0.3, 0.3 ]):
 
 dots = { "a" : 1, "abc" : 1, "ac" : 1, "bc" : 2 }
 make_triple_diagram(dots, figsize = (1,1))
-add_triple_labels(["w_1","w_2","w_3"], pads = [ 0.25, 0.5, 0.55 ])
+add_triple_labels(["u","v","w"])
 plt.tight_layout(pad = 0)
-plt.savefig(fig_dir + "example_triple.pdf")
-
-make_triple_diagram(dots, { "a" : "o", "ac" : "o" }, figsize = (1,1))
-add_triple_labels(["w_1","w_2","w_3"], pads = [ 0.25, 0.5, 0.55 ])
-plt.tight_layout(pad = 0)
-plt.savefig(fig_dir + "example_triple_oo_2.pdf")
-
-make_triple_diagram(dots, { "ac" : "o", "bc" : "o" }, figsize = (1,1))
-add_triple_labels(["w_1","w_2","w_3"], pads = [ 0.25, 0.5, 0.55 ])
-plt.tight_layout(pad = 0)
-plt.savefig(fig_dir + "example_triple_oo_1.pdf")
[…]

### 2020-02-07 collective_sun/sun_perturbation.tex
cleaned up appendix on products of multi-body operators

@@
 
-\newcommand{\FS}{\text{FS}}
-\newcommand{\EQFS}{=_{\text{FS}}}
+\newcommand{\PS}{\text{PS}}
+\newcommand{\EQPS}{=_{\text{PS}}}
 \newcommand{\col}{\underline}
@@
   H_1^{(1)} = \sum_{X,p} v_{Xp} \P_0 X_p \P_0
-  \EQFS \sum_{X,p} v_{Xp} X_1
-  \EQFS \sum_X \bar v_X \col{X},
+  \EQPS \sum_{X,p} v_{Xp} X_1
+  \EQPS \sum_X \bar v_X \col{X},
   \label{eq:H_1_1}
 \end{align}
-where $\EQFS$ denotes equality under a restriction to the fully
+where $\EQPS$ denotes equality under a restriction to the fully
 symmetric manifold $\M_0$, and
@@
   H_1^{(2)}
-  \EQFS \f1{N\p{N-1}} \sum_{X,Y} \sum_{\Delta\ne0}
+  \EQPS \f1{N\p{N-1}} \sum_{X,Y} \sum_{\Delta\ne0}
   \f{\v v_X^\Delta\c\v v_Y^\Delta}{\Delta}
@@
 \end{align}
-where $\EQFS$ denotes equality under a restriction to the fully
+where $\EQPS$ denotes equality under a restriction to the fully
 symmetric manifold $\M_0$; $\v v_X^\Delta$ is the projection of
@@
   H_2^{(1)} = \sum_O \sum_{p<q} w_{Opq} \P_0 O_{pq} \P_0
-  \EQFS \sum_O \sum_{p<q} w_{Opq} O_{1,2}
-  \EQFS \sum_O \bar w_O \col{O},
+  \EQPS \sum_O \sum_{p<q} w_{Opq} O_{1,2}
+  \EQPS \sum_O \bar w_O \col{O},
 \end{align}
@@
 \begin{align}
-  H_2^{(1)} \EQFS \f12 \sum_{\p{X,Y}} \bar w_{\p{X,Y}}
+  H_2^{(1)} \EQPS \f12 \sum_{\p{X,Y}} \bar w_{\p{X,Y}}
   \p{\col{X}\,\col{Y} - \col{XY}}.
@@
 \begin{align}
-  H_2^{(2)} \EQFS W_4 S_\z^4 - W_2 S_\z^2,
-\end{align}
-where $S_\z\equiv\sum_p s_\z^{(p)}$ is a collective SU(2) spin
-operator, $W_k$ are scalar coefficients, and we have neglected scalar
-terms in $H_2^{(2)}$ without physical consequence.  In order to expand
-the coefficients $W_k$, for any ${N \choose 2}$-component vector
+  H_2^{(2)} \simeq_\PS W_4 S_\z^4 - W_2 S_\z^2,
+\end{align}
+where $\simeq_\PS$ denotes equality up to scalar terms and a
+restriction to the permutationally symmetric manifold;
+$S_\z\equiv\sum_p s_\z^{(p)}$ is a collective SU(2) spin operator; and
+$W_k$ are scalar coefficients.  In order to expand the coefficients
+$W_k$, for any ${N \choose 2}$-component vector
 $\v{\m m}=\sum_{p<q}m_{pq}\ket{\set{p,q}}$ we define the $N$-component
@@
   \prod_{j\in\ZZ_p} \sum_{k\in\ZZ_N^{M_j}} w_j\p{k} O_j\p{k}
-  \EQFS \sum_{g\in\G\p{\m M}} \S\p{g} w\p{g} O\p{g},
+  \EQPS \sum_{g\in\G\p{\m M}} \S\p{g} w\p{g} O\p{g},
   \label{eq:sym_prod_group_start}
 \end{align}
-where $\EQFS$ denotes equality up to a restriction to the fully
+where $\EQPS$ denotes equality up to a restriction to the fully
 symmetric manifold, $O\p{g}$ is an operator acquired by assigning
@@
 that account for the number of ways to partition the indices of each
-$w_j,O_j$ into sets of shared indices according to $g_S$ for all
-$j\in S$, and the second product in \eqref{eq:sym_prod_symmetry}
-accounts for the number of ways to permute each set of shared indices
-on every tensor.  Altogether, the symmetry factors $\S\p{g}$ cancel
-out with the $\sim M_j!$ prefactors in \eqref{eq:sym_prod_vec}, so
+$w_j,O_j$ into sets of shared indices as specified by the values of
+$g_S$ for all $j\in S$, and the second product in
+\eqref{eq:sym_prod_symmetry} accounts for the number of ways to
+permute each set of shared indices on every tensor.  Altogether, the
+symmetry factors $\S\p{g}$ cancel out with the $\sim M_j!$ prefactors
+in \eqref{eq:sym_prod_vec}, so
 \begin{align}
   \prod_{j\in\ZZ_p} \sum_{k\in\C_N\p{M_j}} w_j\p{k} O_j\p{k}
-  \EQFS \sum_{g\in\G\p{\m M}} w\p{g} O\p{g}.
+  \EQPS \sum_{g\in\G\p{\m M}} w\p{g} O\p{g}.
   \label{eq:sym_prod_group}
@@
 Diagrams such as \eqref{eq:venn_diagram} will play a major role in the
-calculations in this section.  Here, we first define diagrams as
-formal objects that we can use in mathematical expressions.
-Formalizing diagrams will enable us to perform necessary calculations
-that are otherwise intractable.  We then discuss rules for decomposing
-these diagrams in order to reduce the computational cost of their
-numerical evaluation.
+calculations in this section.  Here, we define these diagrams as
+formal objects that we can use in mathematical expressions, enabling
+us to systematically perform calculations that are otherwise
+intractable.  We then derive rules to ``simplify'' these diagrams in
+such a way as to reduce the computational cost of their numerical
+evaluation.
 
 The simplest diagrams have $g_S$ dots (``$\bullet$'') at the
-intersection of the circles $j\in S$, and are simply equal to the
-associated scalar $w\p{g}$ defined in \eqref{eq:diagram_factors}, e.g.
+intersection of the circles $j\in S\in\PP\p{\ZZ_p}$, and are simply
+equal to the associated scalar $w\p{g}$ defined in
+\eqref{eq:diagram_factors}, e.g.
 \begin{align}
@@
 strategy will be to replace constrained sums that make all indices
-interdependent (such as $\abs{\set{a,b,c,d,e}}=5$ above) by
-unconstrained sums that can be carried out in sequence.  In order to
-represent different constraints, we need to define diagrams in which
-filled dots (``$\bullet$'') may be replaced by empty dots
-(``$\circ$'') or crosses (``$\bm\times$'').
+interdependent (e.g.~$\abs{\set{a,b,c,d,e}}=5$ above) by unconstrained
+sums that can be carried out in sequence.  In order to represent
+different constraints, we need to define diagrams in which filled dots
+(``$\bullet$'') may be replaced by empty dots (``$\circ$'') or crosses
+(``$\bm\times$'').
 
@@
 cost.  We will essentially use two tricks to decompose diagrams:
-eliminating filled dots, and eliminating crosses.
+eliminating filled dots (which will add empty dots and crosses), and
+eliminating crosses (which will move around filled dots).
 
@@
 Given a diagram containing only dots (filled or empty), we can
-decompose any filled dot into an empty dot and a cross.  For example
+decompose any filled dot into an empty dot and a cross.  For example,
 \begin{align}
@@
 \end{align}
-The decomposition in \eqref{eq:example_elim} essentially breaks
-up a sum over $a\in\ZZ_N\setminus\set{b,c}$ into a difference of sums
-over $a\in\ZZ_N$ and $a\in\set{b,c}$.  The sum over $a\in\ZZ_N$ to
-compute $v\p{\circ,b}$ has $\O\p{N^2}$ cost, and can be performed
-prior to evaluating the diagram in \eqref{eq:example_elim_o}.
-The decomposition in \eqref{eq:example_elim} therefore splits
-one $\O\p{N^3}$ into two $\O\p{N^2}$ diagrams.
-
-After splitting a filled dot into an empty dot and a cross, the next
+The decomposition in \eqref{eq:example_elim} essentially breaks up a
+sum over $a\in\ZZ_N\setminus\set{b,c}$ into a difference of sums over
+$a\in\ZZ_N$ and $a\in\set{b,c}$.  The sum over $a\in\ZZ_N$ to compute
+$v\p{\circ,b}$ has $\O\p{N^2}$ cost, and can be performed prior to
+evaluating the diagram in \eqref{eq:example_elim_o}.  The
+decomposition in \eqref{eq:example_elim} therefore splits one
+$\O\p{N^3}$ diagram into two $\O\p{N^2}$ diagrams.
+
+After decomposing a filled dot into an empty dot and a cross, the next
 step is to immediately eliminate the cross.  To eliminate a cross from
@@
   = 2 \diagram{example_elim_x_full}
-  = 2 \sum_{b,c\in\C_N\p{2}} v\p{c,b} w\p{b,c}.
+  = 2 \sum_{\p{b,c}\in\C_N\p{2}} v\p{c,b} w\p{b,c}.
   \label{eq:example_elim_final}
@@
 region with $F_{\t{old}}$ filled dots into a ``new'' region with
-$F_{\t{new}}$ filled dots picks up an overall factor of $F_{\t{new}}$,
-so
+$F_{\t{new}}$ filled dots picks up an overall factor of
+$F_{\t{new}}+1$, so
 \begin{align}
@@
 \end{align}
-where the factor of $1/F_{\t{old}}$ that would be acquired from
-removing a dot from the old region is canceled out by a factor of
-$F_{\t{old}}$ that accounts for the number of ways to move a dot from
-the old region to the new one.
+where the factor of $1/F_{\t{old}}$ that is acquired from removing a
+dot from the old region is canceled out by a factor of $F_{\t{old}}$
+that accounts for the number of ways to move a dot from the old region
+to the new one.
 
@@
   \P_0 \sp{\sum_{p,q} v_p w_q X_p Y_q} \P_0
-  \EQFS \diagram{single_body_0} X_1 Y_2
+  \EQPS \diagram{single_body_0} X_1 Y_2
   + \diagram{single_body_1} X_1 Y_1,
@@
   = \diagram{single_body_0_o} - \diagram{single_body_0_x}
-  = \diagram{single_body_0_oo} - \diagram{single_body_1}
+  = \diagram{single_body_0_o} - \diagram{single_body_1}
   = - \v v \c\v w,
@@
   \sum_{p,q} v_p w_q X_p Y_q
-  \EQFS - \v v\c\v w \p{X_1 Y_2 - X_1 Y_1},
+  \EQPS - \v v\c\v w \p{X_1 Y_2 - X_1 Y_1},
 \end{align}
@@
 \begin{align}
-  \col{X Y} \EQFS N X_1 Y_1,
+  \col{X Y} \EQPS N X_1 Y_1,
   &&
@@
   = \sum_{p=q} X_p Y_q + \sum_{p\ne q} X_p Y_q
-  \EQFS \col{XY} + N\p{N-1} X_1 Y_2,
+  \EQPS \col{XY} + N\p{N-1} X_1 Y_2,
 \end{align}
@@
   \sum_{p,q} v_p w_q X_p Y_q
-  \EQFS - \f{\v v \c\v w}{N\p{N-1}}
+  \EQPS - \f{\v v \c\v w}{N\p{N-1}}
   \p{\col{X}\,\col{Y} - N \col{XY}}.
@@
     v_{k\ell} w_{pq} O_{k\ell} Q_{pq}} \P_0 \\
-  \EQFS \diagram{two_body_0} O_{1,2} Q_{3,4}
+  \EQPS \diagram{two_body_0} O_{1,2} Q_{3,4}
   + \diagram{two_body_1} O_{1,2} Q_{1,3}
@@
 where $k,\ell,p,q$ index individual spins; $O$ and $Q$ are
-permutationally-symmetric two-spin operators; the coefficients
+permutationally-symmetric two-spin operators; and the coefficients
 $v_{k\ell},w_{pq}$ satisfy \eqref{eq:cond_2} with eigenvalue
-$\Delta\ne0$, which implies that they are mean-zero; and the factors
-of $1/4$ and $1/2$ are included to cancel out the symmetry factors
-that appear in the definition of a diagram.  Defining
+$\Delta\ne0$, which implies that they are mean-zero.  Defining
 \begin{align}
@@
   \sum_{k<\ell} \sum_{p<q} v_{k\ell} w_{pq} O_{k\ell} Q_{pq} \\
-  \EQFS - \p{\v v \c \v w - \v{\m v} \c \v{\m w}} O_{1,2} Q_{3,4}
+  \EQPS - \p{\v v \c \v w - \v{\m v} \c \v{\m w}} O_{1,2} Q_{3,4}
   + \p{\v v\c\v w - 2 \v{\m v} \c \v{\m w}} O_{1,2} Q_{1,3}
@@
 \begin{align}
-  \X \EQFS A_6 X^{\otimes 6} + A_4 X^{\otimes 4} + A_2 X^{\otimes 2},
+  \X \EQPS A_6 X^{\otimes 6} + A_4 X^{\otimes 4}
+  + A_2 X^{\otimes 2} + A_0,
   \label{eq:triple_multi}
@@
   A_4 \equiv \diagram{triple_01} + \diagram{triple_1},
+  &&
+  A_0 \equiv \diagram{triple_0111},
   \label{eq:triple_64}
@@
   A_2 = D_2 - 3 D_3,
+  &&
+  A_0 \equiv D_3,
 \end{align}
@@
 \begin{align}
-  \X \EQFS
+  \X \simeq_\PS
   \tilde A_6 \col{X}^6 + \tilde A_4 \col{X}^4 + \tilde A_2 \col{X}^2,
@@
 \end{align}
-where we have neglected scalar terms without physical consequence, and
+where $\simeq_\PS$ denotes equality up to scalar terms and a
+restriction to the permutationally symmetric manifold, and
 \begin{align}
@@
 %   \O_S
-%   \EQFS \sum_{\substack{\t{partitions}\\P~\t{of}~S}}
+%   \EQPS \sum_{\substack{\t{partitions}\\P~\t{of}~S}}
 %   N!_{\abs{P}} O_P,
@@
 % \end{align}
-% where $\EQFS$ denotes equality under a restriction to the fully
+% where $\EQPS$ denotes equality under a restriction to the fully
 % symmetric manifold.  For any multi-local operator $O_S$ indexed by a
@@
 %   N!_{\abs{S}} O_S
-%   &\EQFS \sum_{\substack{\t{partitions}\\P~\t{of}~S}}
+%   &\EQPS \sum_{\substack{\t{partitions}\\P~\t{of}~S}}
 %   \p{-1}^{\abs{S}-\abs{P}} \O_P \times \prod_{p\in P}\p{\abs{p}-1}!
 %   \label{eq:multi_to_collective_full} \\
-%   &\EQFS \f1{\abs{S}!} \sum_{\substack{\t{permutations}\\\Pi~\t{of}~S}}
+%   &\EQPS \f1{\abs{S}!} \sum_{\substack{\t{permutations}\\\Pi~\t{of}~S}}
 %   \sum_{\substack{\t{ordered}\\\t{partitions}\\\tilde P~\t{of}~\Pi}}
@@
 % \begin{align}
-%   \O_{\tilde S} \EQFS
+%   \O_{\tilde S} \EQPS
 %   \sum_{\substack{\t{ordered}\\\t{partitions}\\\tilde P~\t{of}~\tilde S}}
@@
 %   N!_2 X\otimes Y
-%   &\EQFS \col{X}\,\col{Y} - \col{XY}, \\
+%   &\EQPS \col{X}\,\col{Y} - \col{XY}, \\
 %   N!_3 X\otimes Y \otimes Z
-%   &\EQFS \col{X}\,\col{Y}\,\col{Z}
+%   &\EQPS \col{X}\,\col{Y}\,\col{Z}
 %   - \col{X}\,\col{YZ} - \col{Y}\,\col{XZ} - \col{Z}\,\col{XY}
@@
 % \begin{align}
-%   N!_2 X\otimes X \EQFS \col{X}^2 - \col{X^2},
+%   N!_2 X\otimes X \EQPS \col{X}^2 - \col{X^2},
 %   &&
-%   N!_3 X^{\otimes 3} \EQFS \col{X}^3
+%   N!_3 X^{\otimes 3} \EQPS \col{X}^3
 %   - 3 \col{X} \odot \col{X^2} + 2\col{X^3},
@@
 %   N!_4 X^{\otimes 4}
-%   \EQFS \col{X}^4 - 4 \col{X}^2 \odot \col{X^2}
+%   \EQPS \col{X}^4 - 4 \col{X}^2 \odot \col{X^2}
 %   - 2 \col{X} \, \col{X}^2 \, \col{X}
@@
 % \begin{align}
-%   N!_2 X\otimes X \EQFS \col{X}^2 - N,
+%   N!_2 X\otimes X \EQPS \col{X}^2 - N,
 %   &&
-%   N!_3 X^{\otimes 3} \EQFS \col{X}^3 - \p{3N-2} \col{X},
+%   N!_3 X^{\otimes 3} \EQPS \col{X}^3 - \p{3N-2} \col{X},
 % \end{align}
 % \begin{align}
-%   N!_4 X^{\otimes 4} \EQFS \col{X}^4
+%   N!_4 X^{\otimes 4} \EQPS \col{X}^4
 %   - 2\p{3N-4} \col{X}^2 + 3 N \p{N-2}.

### 2020-02-07 collective_sun/sun_perturbation.tex
fixed a couple of minor errors
minor error fix

@@
   \diagram{example_elim}
-  \equiv \sum_{\substack{a,b,c\in\ZZ_N\\\abs{\set{a,b,c}}=2}}
+  \equiv \sum_{\substack{a,b,c\in\ZZ_N\\\abs{\set{a,b,c}}=3}}
   v\p{a,b} w\p{b,c},
@@
 \end{align}
-where the factor of $1/F_{\t{old}}$ that is acquired from removing a
+where the factor of $1/F_{\t{old}}$ that is acquired due to removing a
 dot from the old region is canceled out by a factor of $F_{\t{old}}$
@@
 \begin{align}
-  \diagram{two_body_2} \equiv \sum_{p,q\in\C_N\p{2}} v_{pq} w_{pq}
+  \diagram{two_body_2} \equiv \sum_{\p{p,q}\in\C_N\p{2}} v_{pq} w_{pq}
   = \v{\m v} \c\v{\m w},
@@
 
-Here we simplify a product of three two-body operators of a particular
-form projected onto the permutationally symmetric manifold:
+Here we simplify a product of three two-body operators projected onto
+the permutationally symmetric manifold:
 \begin{align}
@@
 \end{align}
-where we have neglected scalar terms without physical consequence, and
-a diagram with unlabeled circles denotes a sum over all distinct label
-assignments, e.g.
+where a diagram with unlabeled circles denotes a sum over all distinct
+label assignments, e.g.
 \begin{align}

### 2020-02-10 collective_sun/sun_perturbation.tex
simplify the two-body problem in the case of an isotropic system

@@
 
+\newcommand{\DD}{\mathbb{D}}
 \newcommand{\PP}{\mathbb{P}}
+\newcommand{\RR}{\mathbb{R}}
 \renewcommand{\SS}{\mathbb{S}}
@@
 the two-body eigenvalue problem to that of diagonalizing
-$\p{N-1}\times\p{N-1}$ matrices.  Finally, for we discuss the further
-reduction of the two-body problem to that of diagonalizing matrices
-with dimensions $\sim\p{N/2^D}\times\p{N/2^D}$ in the case of an
-isotropic system on a $D$-dimensional periodic lattice, but leave this
-reduction to future work.
+$\p{N-1}\times\p{N-1}$ matrices.  We then reduce the two-body
+eigenvalue problem even further for the case of an isotropic system,
+to that of diagonalizing a single $\sim\p{N/2^D}\times\p{N/2^D}$
+matrix.  Finally, we provide an analytic solution to the two-body
+problem in the case of a fully collective model with uniform
+couplings.
 
@@
 \begin{align}
-  \sum_k \p{h_{pk} w_{kq} + w_{pk} h_{kq}}
-  - \bar h_{pq} w_{pq}
+  \sum_k \p{h_{pk} w_{kq} + w_{pk} h_{kq}} - \bar h_{pq} w_{pq}
   = \Delta w_{pq},
@@
 \end{align}
-which can be written in matrix form as
-\begin{align}
-  \P_{\t{off-diag}}\p{\m h\c\m w + \m w\c\m h} - \bar{\m h} * \m w
-  = \Delta \m w,
+where translational invariance implies that
+$\bar h_{pq}=2h_0-2h_{pq}$, so we can equivalently write
+\begin{align}
+  h_{pq} w_{pq}
+  + \f12 \sum_k \p{h_{pk} w_{kq} + w_{pk} h_{kq}}
+  = \p{\f12 \Delta + h_0} w_{pq},
+\end{align}
+which can also be written in matrix form as
+\begin{align}
+  \m h * \m w + \f12 \P_{\t{off-diag}}\p{\m h\c\m w + \m w\c\m h}
+  = \p{\f12 \Delta + h_0} \m w,
   \label{eq:cond_2_ref}
 \end{align}
-where $\m h$, $\bar{\m h}$, and $\m w$ are matrices of all $h_{pq}$,
-$\bar h_{pq}$ and $w_{pq}$;
+where $\m h$ and $\m w$ are matrices of all $h_{pq}$ and $w_{pq}$;
+$\m h*\m w$ denotes an element-wise (Kronecker) product of $\m h$ and
+$\m w$, i.e.~with matrix elements $h_{pq}w_{pq}$; and
 $\P_{\t{off-diag}}\p{\m m}\equiv \m m-\sum_p\op{p}\m m\op{p}$ denotes
 the projection of $\m m$ onto the space of strictly off-diagonal
-matrices; and $\bar{\m h}*\m w$ denotes an element-wise (Kronecker)
-product of $\bar{\m h}$ and $\m w$, i.e.~with matrix elements
-$\bar h_{pq}w_{pq}$.  In order to find solutions to
-\eqref{eq:cond_2_ref} when the interaction Hamiltonian $H_0$ is
-translationally invariant, in which case $h_{pq}$ and $\bar h_{pq}$
-depend only on the separation $\pm\p{p-q}$, we expand all matrices in
-a basis that the respects translational symmetry.  To this end, for
-all lattice displacements $d\in\ZZ_L^D$ and wavenumbers
-$k\in\ZZ_L^D\times2\pi/L$ we define the {\it cycle matrices}
+matrices.  In order to find solutions to \eqref{eq:cond_2_ref} when
+the interaction Hamiltonian $H_0$ is translationally invariant, in
+which case $h_{pq}$ depends only on the separation $\pm\p{p-q}$, we
+expand all matrices in a basis that the respects translational
+symmetry.  To this end, for all lattice displacements $d\in\ZZ_L^D$
+and wavenumbers $k\in\ZZ_L^D\times2\pi/L$ we define the {\it cycle
+  matrices}
 \begin{align}
@@
   \m h = \sum_{q,d} h_{q+d,q} \op{q+d}{q}
-  = \sum_d h_{d,0} \m m_{d,0},
-  &&
-  \bar{\m h} = \sum_d \bar h_{d,0} \m m_{d,0}.
+  = \sum_d h_{d,0} \m m_{d,0}.
 \end{align}
@@
 \begin{align}
-  \sum_{\substack{c,d\\c+d\ne0}}
+  \sum_{d\ne0} h_{d,0} \tilde w_{dk} \m m_{dk}
+  + \f12 \sum_{\substack{c,d\\c+d\ne0}}
   h_{c,0} \tilde w_{dk} \p{1 + e^{-ic\c k}} \m m_{c+d,k}
-  - \sum_{d\ne0} \bar h_{d,0} \tilde w_{dk} \m m_{dk}
-  = \Delta \sum_{d\ne0} \tilde w_{dk} \m m_{dk},
+  = \p{\f12\Delta + h_0} \sum_{d\ne0} \tilde w_{dk} \m m_{dk},
 \end{align}
@@
 \begin{align}
-  \sum_{c,d\ne0} h_{d-c,0} \tilde w_{ck}
+  \sum_{d\ne0} h_{d,0} \tilde w_{dk} \m m_{dk}
+  + \f12 \sum_{c,d\ne0} h_{d-c,0} \tilde w_{ck}
   \p{1 + e^{-i\p{d-c}\c k}} \m m_{dk}
-  - \sum_{d\ne0} \bar h_{d,0} \tilde w_{dk} \m m_{dk}
-  = \Delta \sum_{d\ne0} \tilde w_{dk} \m m_{dk},
+  = \p{\f12\Delta + h_0} \sum_{d\ne0} \tilde w_{dk} \m m_{dk},
   \label{eq:cond_2_cycle}
@@
   \ut{\check{\m h}}_k
-  \equiv \sum_{c,d\ne0} h_{d-c,0} \p{1+e^{-i\p{d-c}\c k}} \op{d}{c}
-  - \sum_{d\ne0} \bar h_{d,0} \op{d},
+  \equiv \sum_{d\ne0} h_{d,0} \op{d}
+  + \f12 \sum_{c,d\ne0} h_{d-c,0} \p{1+e^{-i\p{d-c}\c k}} \op{d}{c},
   &&
@@
 \begin{align}
-  \ut{\check{\m h}}_k \c \ut{\v{\m w}}_k = \Delta \ut{\v{\m w}}_k.
-  \label{eq:cond_2_square}
-\end{align}
-The appearance of a wavenumber index $k$ in \eqref{eq:cond_2_square}
-implies that all fixed-$k$ components of a general coefficient matrix
-$\m w$ must simultaneously satisfy \eqref{eq:cond_2_square} in order
-to generate states of fixed excitation energy $\Delta$ when the
-corresponding perturbation $\V_2^\Delta$ acts on a permutationally
-symmetric state $\ket\psi\in\M_0$.  The decomposition of an arbitrary
-coefficient vector $\v{\m w}=\sum_\Delta\v{\m w}_\Delta$ into
-coefficient vectors $\v{\m w}_\Delta$ that generate a
-fixed-excitation-energy perturbation $\V_2^\Delta$ is then performed
-by projecting each $k$-component of $\v{\m w}$ onto the corresponding
-$\Delta$-eigenspace of $\ut{\check{\m h}}_k$.  If the perturbation
-$\V_2$ is translationally invariant, then the $k$-components of
-$\v{\m w}$ for all $k\ne 0$ all vanish, so we only need to consider
-the case of $k=0$.
-
-We can reduce the eigenvalue problem in \eqref{eq:cond_2_square} even
-further in the case of an isotropic system, in which the couplings
-$h_{pq}$ depend only on the distance $\abs{p-q}$.  On a $D$
-dimensional lattice, isotropy implies one reflection symmetry for each
-lattice axis, which reduces the degrees of freedom in of
-$\ut{\check{\m h}}_k$ from $\sim N\times N$ to
-$\sim\p{N/2^D}\times\p{N/2^D}$.  We leave this reduction to future
-work.
+  \ut{\check{\m h}}_k \c \ut{\v{\m w}}_k
+  = \p{\f12\Delta + h_0} \ut{\v{\m w}}_k.
+  \label{eq:cond_2_cycle_block}
+\end{align}
+The appearance of a wavenumber index $k$ in
+\eqref{eq:cond_2_cycle_block} implies that all fixed-$k$ components of
+a general coefficient matrix $\m w$ must simultaneously satisfy
+\eqref{eq:cond_2_cycle_block} in order to generate states of fixed
+excitation energy $\Delta$ when the corresponding perturbation
+$\V_2^\Delta$ acts on a permutationally symmetric state
+$\ket\psi\in\M_0$.  The decomposition of an arbitrary coefficient
+vector $\v{\m w}=\sum_\Delta\v{\m w}_\Delta$ into coefficient vectors
+$\v{\m w}_\Delta$ that generate a fixed-excitation-energy perturbation
+$\V_2^\Delta$ is then performed by projecting each $k$-component of
+$\v{\m w}$ onto the corresponding $\Delta$-eigenspace of
+$\ut{\check{\m h}}_k$.  If the perturbation $\V_2$ is translationally
+invariant, then $\ut{\v{\m w}}_k=0$ for all $k\ne0$, so we only need
+to consider the case of $k=0$:
+\begin{align}
+  \ut{\check{\m h}}_0 \c \ut{\v{\m w}}_0
+  = \p{\f12\Delta + h_0} \ut{\v{\m w}}_0,
+  &&
+  \ut{\check{\m h}}_0
+  = \sum_{d\ne0} h_{d,0} \op{d}
+  + \sum_{c,d\ne0} h_{cd} \op{c}{d},
+  &&
+  \ut{\v{\m w}}_0 \equiv \sum_{d\ne0} \tilde w_{d,0} \ket{d}.
+  \label{eq:cond_2_cycle_0}
+\end{align}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Isotropic systems}
+
+If the interaction Hamiltonian $H_0$ and two-body perturbation $\V_2$
+are both translationally invariant isotropic, then we can reduce the
+eigenvalue problem \eqref{eq:cond_2_cycle_0} even further by making
+use of the fact that the couplings $h_{pq}$ depend only on the
+distance $\abs{p-q}$.  To this end, we define vectors $\ket{a}$
+labeled by distances $a\in\RR$ from the origin:
+\begin{align}
+  \ket{a} \equiv \f1{\sqrt{N_a}}
+  \sum_{\substack{d\in\ZZ_L^D\\\abs{d}=a}} \ket{d},
+\end{align}
+where the norm $\abs{d}$ of a vector $d$ on a periodic lattice is
+implicitly understood to mean the smallest Euclidean distance of $d$
+from the origin; and $N_a$ is a normalization factor equal to the
+number of displacements $d\in\ZZ_L^D$ with $\abs{d}=a$, such that
+$\bk{a|a}=1$.  Defining the set of distances
+\begin{align}
+  \DD_{LD} \equiv \set{ \abs{d} : d \in \ZZ_L^D } \setminus \set{0},
+\end{align}
+we can expand
+\begin{align}
+  \ut{\check{\m h}}_0
+  = \sum_{a\in\DD_{LD}} \tilde h_a \op{a}
+  + \sum_{a,b\in\DD_{LD}} \tilde h_{ab} \op{a}{b},
+  &&
+  \ut{\v{\m w}}_0
+  \equiv \sum_{a\in\DD_{LD}} \tilde w_a \sqrt{N_a} \ket{a},
+\end{align}
+where $\tilde h_a=h_{d,0}$ and
+$\tilde w_a=\tilde w_{d,0}=\tr\p{\m m_{d,0}^\dag \m w}/N$ for all
+$\abs{d}=a$, and
+\begin{align}
+  \tilde h_{ab} \equiv \f1{\sqrt{N_aN_b}}
+  \sum_{\substack{c,d\in\ZZ_L^D\\\abs{c}=a\\\abs{d}=b}} h_{cd}.
+\end{align}
+By rough accounting, isotropy in $D$ dimensions implies one reflection
+symmetry for each of $D$ axes, so the matrix $\ut{\check{\m h}}_0$ has
+dimensions $\sim\p{N/2^D}\times\p{N/2^D}$.
 

### 2020-02-10 collective_sun/sun_perturbation.tex
remove restriction of zero-mena coefficients in the products of two single- and two-body operators

@@
 where the product $\P_0 \p{\V_1^\Delta}^2 \P_0$ is worked out in
-Appendix \ref{sec:PXYP}.  In total, we find
+Appendix \ref{sec:PXYP}.  In total, using the fact that
+$\v v_X^\Delta$ is mean-zero for all $\Delta\ne0$, we find
 \begin{align}
@@
   \p{\col{X}\,\col{Y} - N \col{XY}},
+  &&
+  \col{Z} \equiv \sum_p Z_p.
   \label{eq:H_1_2}
@@
 symmetric manifold $\M_0$; $\v v_X^\Delta$ is the projection of
-$\v v_X$ onto the $\Delta$-eigenspace of the matrix $\m h-\diag\v h$;
-and for any single-spin operator $Z$, we define the collective
-operator
-\begin{align}
-  \col{Z} \equiv \sum_p Z_p.
-\end{align}
+$\v v_X$ onto the $\Delta$-eigenspace of the matrix $\m h-\diag\v h$.
 
@@
 $W_k$ are scalar coefficients.  In order to expand the coefficients
-$W_k$, for any ${N \choose 2}$-component vector
-$\v{\m m}=\sum_{p<q}m_{pq}\ket{\set{p,q}}$ we define the $N$-component
-vector
-\begin{align}
-  \v m \equiv \sum_{p,q} m_{pq} \ket{p},
-\end{align}
-where we define $m_{pq}$ for all $p,q$ (as opposed to only $p<q$) by
-enforcing $m_{pq}=m_{qp}$ and $m_{pp}=0$.  We then denote the
-projection of $\v{\m m}$ onto the $\Delta$-eigenspace of the matrix
-$\check{\m h}$ defined in \eqref{eq:h_super_mat} by $\v{\m m}_\Delta$,
-which allows us to concisely write
-\begin{align}
-  W_4
-  \equiv \sum_{\Delta\ne0} \f1\Delta \f{\v w_\Delta\c\v w_\Delta
-    - \v{\m w}_\Delta\c\v{\m w}_\Delta}{N!_4},
-\end{align}
-and
-\begin{align}
-  W_2
-  \equiv \sum_{\Delta\ne0} \f1{\Delta} \sp{\f{\v w_\Delta\c\v w_\Delta
-      - 2\v{\m w}_\Delta\c\v{\m w}_\Delta}{4N!_2}
-    + \f{\p{3N-4}\p{\v w_\Delta\c\v w_\Delta
-        - \v{\m w}_\Delta\c\v{\m w}_\Delta}}{2N!_4}},
+$W_k$, we define
+\begin{align}
+  \v{\m w}=\sum_{p<q} w_{pq} \ket{\set{p,q}},
+  &&
+  \v w \equiv \sum_{p,q} w_{pq} \ket{p},
+\end{align}
+where $w_{pq}$ is defined for all $p,q$ (as opposed to only $p<q$) by
+enforcing $w_{pq}=w_{qp}$ and $w_{pp}=0$.  The vector $\v w$ is the
+sum of all columns of the matrix
+$\m w\equiv\sum_{p,q}w_{pq}\op{p}{q}$.  We then denote the projection
+of $\v{\m w}$ onto the $\Delta$-eigenspace of the matrix
+$\check{\m h}$ defined in \eqref{eq:h_super_mat} by $\v{\m w}_\Delta$,
+and define
+\begin{align}
+  w_{\t{columns}}^{(2)} \equiv \sum_{\Delta\ne0}
+  \f{\v w_\Delta\c\v w_\Delta}{\Delta},
+  &&
+  w_{\t{pairs}}^{(2)} \equiv \sum_{\Delta\ne0}
+  \f{\v{\m w}_\Delta\c\v{\m w}_\Delta}{\Delta},
+\end{align}
+in terms of which
+\begin{align}
+  W_4 \equiv \f{w_{\t{columns}}^{(2)} - w_{\t{pairs}}^{(2)}}{N!_4},
+  &&
+  W_2 \equiv \f{w_{\t{columns}}^{(2)} - 2 w_{\t{pairs}}^{(2)}}
+  {4N!_2} + \f12 \p{3N-4} W_4,
 \end{align}
@@
 
-Calculating the second-order effective Hamiltonian $H_1^{(2)}$ induced
-on the permutationally symmetric manifold $\M_0$ by the perturbation
-$\V_1$ in \eqref{eq:perturbations} requires us to simplify a product
-of the form
+Here we simplify the product
 \begin{align}
@@
 \end{align}
-where $p,q$ index individual spins; $X,Y$ are single-spin operators;
-and the coefficients $v_p,w_q$ satisfy \eqref{eq:cond_1} with
-eigenvalue $\Delta\ne0$, which implies that they are mean-zero.
-Defining $\v m \equiv \sum_p m_p \ket{p}$ for $m\in\set{v,w}$, we can
-simplify
+where $\P_0$ is a projector onto the permutationally symmetric
+manifold $\M_0$; $p,q$ index individual spins; $v_p,w_p$ are scalars;
+and $X,Y$ are single-spin operators.  This product appears, for
+example, in the calculation of the second-order effective Hamiltonian
+$H_1^{(2)}$ induced on the permutationally symmetric manifold $\M_0$
+by the single-body perturbation $\V_1$ in \eqref{eq:perturbations}.
+Defining
+\begin{align}
+  \v m \equiv \sum_p m_p \ket{p},
+  &&
+  \col{m} \equiv \sum_p m_p,
+\end{align}
+for $m\in\set{v,w}$, we can simplify
 \begin{align}
@@
   = \diagram{single_body_0_o} - \diagram{single_body_0_x}
-  = \diagram{single_body_0_o} - \diagram{single_body_1}
-  = - \v v \c\v w,
+  = \diagram{single_body_0_oo} - \diagram{single_body_1}
+  = \col{u}\,\col{v} - \v v \c\v w,
 \end{align}
@@
   \sum_{p,q} v_p w_q X_p Y_q
-  \EQPS - \v v\c\v w \p{X_1 Y_2 - X_1 Y_1},
+  \EQPS \col{u}\,\col{v} X_1 Y_2 - \v v\c\v w \p{X_1 Y_2 - X_1 Y_1},
 \end{align}
@@
   &&
-  \col{X}\,\col{Y}
-  = \sum_{p=q} X_p Y_q + \sum_{p\ne q} X_p Y_q
-  \EQPS \col{XY} + N\p{N-1} X_1 Y_2,
+  \col{X}\,\col{Y} \EQPS \col{XY} + N\p{N-1} X_1 Y_2,
 \end{align}
@@
   \sum_{p,q} v_p w_q X_p Y_q
-  \EQPS - \f{\v v \c\v w}{N\p{N-1}}
-  \p{\col{X}\,\col{Y} - N \col{XY}}.
+  \EQPS \f{\col{u}\,\col{v}}{N\p{N-1}}
+  \p{\col{X}\,\col{Y} - \col{XY}}
+  - \f{\v v \c\v w}{N\p{N-1}}
+  \p{\col{X}\,\col{Y} - N\col{XY}}.
 \end{align}
@@
 
-Calculating the second-order effective Hamiltonian $H_2^{(2)}$ induced
-on the permutationally symmetric manifold $\M_0$ by the perturbation
-$\V_2$ in \eqref{eq:perturbations} requires us to simplify a product
-of the form
+Here we simplify the product
 \begin{multline}
@@
 \end{multline}
-where $k,\ell,p,q$ index individual spins; $O$ and $Q$ are
-permutationally-symmetric two-spin operators; and the coefficients
-$v_{k\ell},w_{pq}$ satisfy \eqref{eq:cond_2} with eigenvalue
-$\Delta\ne0$, which implies that they are mean-zero.  Defining
+where $\P_0$ is a projector onto the permutationally symmetric
+manifold $\M_0$; $k,\ell,p,q$ index individual spins;
+$v_{k\ell},w_{pq}$ are scalar coefficients with $m_{pq}=m_{qp}$ and
+$m_{pp}=0$ for each of $m\in\set{v,w}$; and $O,Q$ are
+permutationally-symmetric two-spin operators.  This product appears,
+for example, in the calculation of the second-order effective
+Hamiltonian $H_2^{(2)}$ induced on the permutationally symmetric
+manifold $\M_0$ by the two-body perturbation $\V_2$ in
+\eqref{eq:perturbations} in \eqref{eq:perturbations}.  Defining
 \begin{align}
@@
   m_p \equiv \sum_q m_{pq},
+  &&
+  \col{m} \equiv \sum_{p<q} m_{pq},
 \end{align}
@@
 where the middle diagram in \eqref{eq:mid_zero} contains an empty sum
-for the cross, and is therefore equal to zero.  Similarly neglecting
-diagrams with empty sums and additionally using the fact that
-$v_{pq},w_{pq}$ are mean-zero, we can work out
+for the cross, and is therefore equal to zero.  Finally, with a bit of
+more work (or a computer) we can work out that
 \begin{align}
   \diagram{two_body_0}
-  &= \f12 \sp{\diagram{two_body_0_o}}
-  - \f12 \sp{\diagram{two_body_0_x}} \\
-  &= \f12 \sp{\diagram{two_body_0_oo} - \diagram{two_body_0_ox}}
-  - \f12 \sp{\diagram{two_body_1}} \\
-  &= \f12 \sp{-\diagram{two_body_1_o}}
-  - \f12 \p{\v v\c\v w - 2\v{\m v}\c\v{\m w}} \\
-  &= -\v v\c\v w + \v{\m v}\c\v{\m w}.
-\end{align}
-Altogether, we thus have that
+  = \diagram{two_body_0_oooo} - \diagram{two_body_1_ooo}
+  + \diagram{two_body_2_oo}
+  = \col{v}\,\col{w} -\v v\c\v w + \v{\m v}\c\v{\m w}.
+\end{align}
+Altogether,
 \begin{multline}
   \sum_{k<\ell} \sum_{p<q} v_{k\ell} w_{pq} O_{k\ell} Q_{pq} \\
-  \EQPS - \p{\v v \c \v w - \v{\m v} \c \v{\m w}} O_{1,2} Q_{3,4}
+  \EQPS \p{\col{v}\,\col{w} -\v v\c\v w + \v{\m v}\c\v{\m w}}
+  O_{1,2} Q_{3,4}
   + \p{\v v\c\v w - 2 \v{\m v} \c \v{\m w}} O_{1,2} Q_{1,3}
@@
 \end{multline}
+In the special case that $O=Q=X\otimes X$ with $X^2=1$, we can further
+simplify
+\begin{align}
+  \sum_{k<\ell} \sum_{p<q} v_{k\ell} w_{pq} X_k X_\ell X_p X_q
+  \simeq_\PS
+  \p{\col{v}\,\col{w} -\v v\c\v w + \v{\m v}\c\v{\m w}} X^{\otimes 4}
+  + \p{\v v\c\v w - 2 \v{\m v} \c \v{\m w}} X^{\otimes 2},
+\end{align}
+where $\simeq_\PS$ denotes equality up to scalar terms and a
+restriction to the permutationally symmetric manifold.  In terms of
+collective operators $\col{X} \equiv \sum_p X_p$, we then have
+\begin{multline}
+  \sum_{k<\ell} \sum_{p<q} v_{k\ell} w_{pq} X_k X_\ell X_p X_q \\
+  \simeq_\PS
+  \f{\col{v}\,\col{w} -\v v\c\v w + \v{\m v}\c\v{\m w}}{N!_4}
+  \sp{\col{X}^4 - 2\p{3N-4} \col{X}^2}
+  + \f{\p{\v v\c\v w - 2 \v{\m v} \c \v{\m w}}}{N!_2}\, \col{X}^2,
+\end{multline}
+where
+\begin{align}
+  N!_k \equiv \prod_{j=0}^{k-1} \p{N-j} = \f{N!}{\p{N-k}!}
+\end{align}
+denotes a falling factorial.
 
@@
 
-Here we simplify a product of three two-body operators projected onto
-the permutationally symmetric manifold:
+Here we simplify the product
 \begin{align}
@@
 \end{align}
-where $k,\ell,p,q,r,s$ index individual spins; $X$ is a single-spin
-operator; and $u_{k\ell},v_{pq},w_{rs}$ are arbitrary scalar
+where $\P_0$ is a projector onto the permutationally symmetric
+manifold $\M_0$; $k,\ell,p,q,r,s$ index individual spins; $X$ is a
+single-spin operator; and $u_{k\ell},v_{pq},w_{rs}$ are scalar
 coefficients with $m_{pq}=m_{qp}$ and $m_{pp}=0$ for each of

### 2020-02-10 collective_sun/scripts/make_diagrams.py

@@ -351,6 +351,12 @@ for assignment in [ { "a" : 2, "b" : 2, "ab" : 0 },
     tag = assignment["ab"]
     plt.savefig(fig_dir + f"two_body_{tag}.pdf")
 
+dots = { "ab" : 2 }
+make_double_diagram(dots, { "ab" : "oo" })
+add_double_labels(["v","w"])
+plt.tight_layout(pad = 0.05)
+plt.savefig(fig_dir + f"two_body_2_oo.pdf")
+
 dots = { "a" : 1, "b" : 1, "ab" : 1 }
 make_double_diagram(dots, { "a" : "o" })
 add_double_labels(["v","w"])
@@ -372,6 +378,11 @@ add_double_labels(["v","w"])
 plt.tight_layout(pad = 0.05)
 plt.savefig(fig_dir + f"two_body_1_ox.pdf")
 
+make_double_diagram(dots, { "a" : "o", "b" : "o", "ab" : "o" })
+add_double_labels(["v","w"])
+plt.tight_layout(pad = 0.05)
+plt.savefig(fig_dir + f"two_body_1_ooo.pdf")
+
 dots = { "a" : 2, "b" : 2 }
 make_double_diagram(dots, { "a" : "o" })
 add_double_labels(["v","w"])
@@ -383,15 +394,10 @@ add_double_labels(["v","w"])
 plt.tight_layout(pad = 0.05)
 plt.savefig(fig_dir + f"two_body_0_x.pdf")
 
-make_double_diagram(dots, { "a" : "oo" })
+make_double_diagram(dots, { "a" : "oo", "b" : "oo" })
 add_double_labels(["v","w"])
 plt.tight_layout(pad = 0.05)
-plt.savefig(fig_dir + f"two_body_0_oo.pdf")
-
-make_double_diagram(dots, { "a" : "ox" })
-add_double_labels(["v","w"])
-plt.tight_layout(pad = 0.05)
-plt.savefig(fig_dir + f"two_body_0_ox.pdf")
+plt.savefig(fig_dir + f"two_body_0_oooo.pdf")
 
 plt.close("all")
 

### 2020-02-11 collective_sun/sun_perturbation.tex
added scalar component of operator products for completion / debugging purposes

@@
   \sum_{k<\ell} \sum_{p<q} v_{k\ell} w_{pq} X_k X_\ell X_p X_q
-  \simeq_\PS
+  \EQPS
   \p{\col{v}\,\col{w} -\v v\c\v w + \v{\m v}\c\v{\m w}} X^{\otimes 4}
-  + \p{\v v\c\v w - 2 \v{\m v} \c \v{\m w}} X^{\otimes 2},
-\end{align}
-where $\simeq_\PS$ denotes equality up to scalar terms and a
-restriction to the permutationally symmetric manifold.  In terms of
-collective operators $\col{X} \equiv \sum_p X_p$, we then have
+  + \p{\v v\c\v w - 2 \v{\m v} \c \v{\m w}} X^{\otimes 2}
+  + \v{\m v}\c\v{\m w}.
+\end{align}
+In terms of collective operators $\col{X} \equiv \sum_p X_p$, we then
+have
 \begin{multline}
   \sum_{k<\ell} \sum_{p<q} v_{k\ell} w_{pq} X_k X_\ell X_p X_q \\
-  \simeq_\PS
+  \EQPS
   \f{\col{v}\,\col{w} -\v v\c\v w + \v{\m v}\c\v{\m w}}{N!_4}
   \sp{\col{X}^4 - 2\p{3N-4} \col{X}^2}
-  + \f{\p{\v v\c\v w - 2 \v{\m v} \c \v{\m w}}}{N!_2}\, \col{X}^2,
+  + \f{\p{\v v\c\v w - 2 \v{\m v} \c \v{\m w}}}{N!_2}\, \col{X}^2 \\
+  + \f{3\,\col{v}\,\col{w} - N \v v\c\v w
+    + N\p{N-2}\v{\m v}\c\v{\m w}}{\p{N-1}\p{N-3}},
 \end{multline}
@@
 \end{align}
-The first diagram in each of $A_6,A_4,A_2$ as written in
+The first diagram in each of $A_6,A_4,A_2,A_0$ as written in
 \eqref{eq:triple_64}, \eqref{eq:triple_2} thus has only one assignment
@@
 Having identified diagrammatic representations of the coefficients
-$A_6,A_4,A_2$, we now need to simplify all diagrams to reduce their
-computational complexity, as discussed in Section \ref{sec:diagrams}.
-Simplifying all these diagrams by hand is tedious, but the process of
-eliminating filled dots and crosses from a diagram is algorithmic
-enough to execute on a computer, doing which we find
+$A_6,A_4,A_2,A_0$, we now need to simplify all diagrams to reduce
+their computational complexity, as discussed in Section
+\ref{sec:diagrams}.  Simplifying all these diagrams by hand is
+tedious, but the process of eliminating filled dots and crosses from a
+diagram is algorithmic enough to execute on a computer, doing which we
+find
 \begin{align}
@@
 \begin{align}
-  \X \simeq_\PS
-  \tilde A_6 \col{X}^6 + \tilde A_4 \col{X}^4 + \tilde A_2 \col{X}^2,
+  \X \EQPS
+  \tilde A_6 \col{X}^6 + \tilde A_4 \col{X}^4
+  + \tilde A_2 \col{X}^2 + \tilde A_0,
   \label{eq:triple_col}
 \end{align}
-where $\simeq_\PS$ denotes equality up to scalar terms and a
-restriction to the permutationally symmetric manifold, and
+where
 \begin{align}
@@
   \tilde A_2 \equiv \f{A_2}{N!_2} - 2\p{3N-4} \f{A_4}{N!_4}
-  + \p{45N^2-210N+184} \f{A_6}{N!_6}.
+  + \p{45N^2-210N+184} \f{A_6}{N!_6},
+\end{align}
+\begin{align}
+  \tilde A_0 \equiv A_0 - \f{A_2}{N-1}
+  + \f{3A_4}{\p{N-1}\p{N-3}}
+  - \f{15A_6}{\p{N-1}\p{N-3}\p{N-5}}.
 \end{align}

### 2020-02-12 collective_sun/sun_perturbation.tex
changed 'fully symmetric' to 'permutationally symmetric'; other minor language changes

@@
 \end{align}
-where $\EQPS$ denotes equality under a restriction to the fully
-symmetric manifold $\M_0$, and
+where $\EQPS$ denotes equality under a restriction to the
+permutationally symmetric manifold $\M_0$, and
 \begin{align}
@@
 \end{align}
-where $\EQPS$ denotes equality under a restriction to the fully
-symmetric manifold $\M_0$; $\v v_X^\Delta$ is the projection of
-$\v v_X$ onto the $\Delta$-eigenspace of the matrix $\m h-\diag\v h$.
+where $\EQPS$ denotes equality under a restriction to the
+permutationally symmetric manifold $\M_0$; $\v v_X^\Delta$ is the
+projection of $\v v_X$ onto the $\Delta$-eigenspace of the matrix
+$\m h-\diag\v h$.
 
@@
 
-The first-order effective Hamiltonian $H_2^{(1)}$ induced on the fully
-symmetric manifold $\M_0$ by the two-body perturbation $\V_2$ in
-\eqref{eq:perturbations} is simply the restriction of $\V_2$ onto
-$\M_0$:
+The first-order effective Hamiltonian $H_2^{(1)}$ induced on the
+permutationally symmetric manifold $\M_0$ by the two-body perturbation
+$\V_2$ in \eqref{eq:perturbations} is simply the restriction of $\V_2$
+onto $\M_0$:
 \begin{align}
@@
 \end{align}
-where $\EQPS$ denotes equality up to a restriction to the fully
-symmetric manifold, $O\p{g}$ is an operator acquired by assigning
-indices to the tensors $O_j$ in a manner consistent with $g$; $w\p{g}$
-is a scalar acquired by summing over all indices assigned to the
-tensors $w_j$ according to $g$; and $\S\p{g}$ is a symmetry factor
-accounting for the number of equivalent ways to assign indices
+where $\EQPS$ denotes equality up to a restriction to the
+permutationally symmetric manifold; $O\p{g}$ is an operator acquired
+by assigning indices to the tensors $O_j$ in a manner consistent with
+$g$; $w\p{g}$ is a scalar acquired by summing over all indices
+assigned to the tensors $w_j$ according to $g$; and $\S\p{g}$ is a
+symmetry factor accounting for the number of ways to assign indices
 according to $g$.
@@
 
-The simplest diagrams have $g_S$ dots (``$\bullet$'') at the
+The simplest diagrams have $g_S$ filled dots (``$\bullet$'') at the
 intersection of the circles $j\in S\in\PP\p{\ZZ_p}$, and are simply
@@
       \abs{\set{a,b,c,d,e}}=5}}
-  w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}.
-\end{align}
-Such a diagram with $\abs{g}$ dots nominally takes $\O\p{N^{\abs{g}}}$
-time to compute.  In practice, this computational cost can be
-prohibitive for performing numerical simulations of systems with any
-appreciable size.  We therefore wish to simplify diagrams in order to
-reduce their computational cost as much as possible.  Our general
-strategy will be to replace constrained sums that make all indices
-interdependent (e.g.~$\abs{\set{a,b,c,d,e}}=5$ above) by unconstrained
-sums that can be carried out in sequence.  In order to represent
+  w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e},
+\end{align}
+where the restriction $\abs{\set{a,b,c,d,e}}=5$ denotes that all of
+$a,b,c,d,e$ are distinct.  Such a diagram with $\abs{g}$ dots
+nominally takes $\O\p{N^{\abs{g}}}$ time to compute.  In practice,
+this computational cost can be prohibitive for performing numerical
+simulations of systems with any appreciable size.  We therefore wish
+to simplify diagrams in order to reduce their computational cost as
+much as possible.  Our general strategy will be to replace constrained
+sums that make all indices interdependent by unconstrained sums that
+can be carried out independently, in sequence.  In order to represent
 different constraints, we need to define diagrams in which filled dots

### 2020-02-28 collective_sun/sun_perturbation.tex
minor change in notation for multinomial coefficient
minor changes to text/naming

@@
 
-We first simplify
+We first consider the single-body case and simplify
 \begin{align}
@@
 \begin{align}
-  \Pi_{p k_j} O_k \ket\psi = O_{k_{jp}}\ket\psi,
+  \Pi_{p k_j} O_k \ket\psi
+  = \Pi_{p k_j} O_k \Pi_{p k_j} \ket\psi
+  = O_{k_{jp}}\ket\psi,
 \end{align}
@@
 use of the fact that the couplings $h_{pq}$ depend only on the
-distance $\abs{p-q}$.  To this end, we define vectors $\ket{a}$
-labeled by distances $a\in\RR$ from the origin:
+distance $\abs{p-q}$.  To this end, we define uniform superpositions
+of all displacements $d\in\ZZ_L^D$ that are a distance $a\in\RR$ from
+the origin:
 \begin{align}
@@
 from the origin; and $N_a$ is a normalization factor equal to the
-number of displacements $d\in\ZZ_L^D$ with $\abs{d}=a$, such that
+number of displacements $d\in\ZZ_L^D$ with $\abs{d}=a$, so that
 $\bk{a|a}=1$.  Defining the set of distances
@@
 \label{sec:sym_prod}
+
+\red{TODO: Discuss motivation, why we care about these products
+  (answer: they appear in perturbation theory, as well as the
+  calculation of matrix elements within the outer-most ``shells'' of
+  fixed net spin).  Give an example of the diagram expansions we're
+  after.}
 
@@
 where $g_S$ is determined the number of dots at the intersection of
-circles $j\in S$.  We denote an assignment of distinct indices
-according to choices of $g_S$ for all $S\in\PP\p{\ZZ_p}$ by $g$, where
-$\PP\p{\ZZ_p}$ is the power set (i.e.~set of all subsets) of $\ZZ_p$.
-Given a vector $\m M$ of the dimensions $M_j$, we then denote the set
-of all valid index assignments $g$ by $\G\p{\m M}$.  For consistency,
-all valid index assignments $g\in\G\p{\m M}$ must assign exactly $M_j$
-indices to the dimension-$M_j$ tensor $w_j$, i.e.~
+circles $\set{w_j:j\in S}$.  We denote an assignment of distinct
+indices according to choices of $g_S$ for all $S\in\PP\p{\ZZ_p}$ by
+q$g$, where $\PP\p{\ZZ_p}$ is the power set (i.e.~set of all subsets)
+of $\ZZ_p$.  Given a vector $\m M$ of the dimensions $M_j$, we then
+denote the set of all valid index assignments $g$ by $\G\p{\m M}$.
+For consistency, all valid index assignments $g\in\G\p{\m M}$ must
+assign exactly $M_j$ indices to the dimension-$M_j$ tensor $w_j$,
+i.e.~
 \begin{align}
@@
 form in \eqref{eq:venn_diagram} with $p=\dim\p{\m M}$ circles and a
-total of $M_j$ dots within circle $j$.
+total of $M_j$ dots within circle $w_j$.
 
@@
 \begin{align}
-  \S\p{g} \equiv \sp{\prod_{j\in\ZZ_p} { M_j \choose g_S : j\in S }}
+  \S\p{g} \equiv \sp{\prod_{j\in\ZZ_p} \C\p{ g_S : j\in S }}
   \sp{\prod_{S\in\PP\p{\ZZ_p}} \p{g_S!}^\abs{S}}
@@
 \begin{align}
-  { M \choose a_1, a_2, \cdots, a_m }
-  \equiv \f{M!}{a_1!a_2!\cdots a_m!}
+  \C\p{a_1,a_2,\cdots,a_m}
+  \equiv \f{\p{a_1+a_2+\cdots+a_m}!}{a_1!a_2!\cdots a_m!}
 \end{align}
@@
 \begin{align}
-  A_6 = D_0 - D_1 + D_2 - D_3,
-  &&
-  A_4 = D_1 - 2 D_2 + 3 D_3,
-  &&
-  A_2 = D_2 - 3 D_3,
-  &&
-  A_0 \equiv D_3,
+  A_6 = D_6 - D_4 + D_2 - D_0,
+  &&
+  A_4 = D_4 - 2 D_2 + 3 D_0,
+  &&
+  A_2 = D_2 - 3 D_0,
+  &&
+  A_0 \equiv D_0,
 \end{align}
@@
 \begin{align}
-  D_0 \equiv \diagram{triple_0_o},
-  &&
-  D_1 \equiv \diagram{triple_01_o} - 2 \diagram{triple_1_o},
-  &&
-  D_3 \equiv \diagram{triple_0111_o},
+  D_6 \equiv \diagram{triple_0_o},
+  &&
+  D_4 \equiv \diagram{triple_01_o} - 2 \diagram{triple_1_o},
+  &&
+  D_0 \equiv \diagram{triple_0111_o},
 \end{align}
@@
 \begin{align}
-  D_0 = \col{u}\,\col{v}\,\col{w},
-  &&
-  D_1 = \col{u}\,\v v\c\v w + \col{v}\,\v w\c\v u
+  D_6 = \col{u}\,\col{v}\,\col{w},
+  &&
+  D_4 = \col{u}\,\v v\c\v w + \col{v}\,\v w\c\v u
   + \col{w}\,\v u\c\v v - 2 \sum_p u_p v_p w_p,
   &&
-  D_3 = \tr\p{\m u \c \m v \c \m w},
+  D_0 = \tr\p{\m u \c \m v \c \m w},
 \end{align}

### 2020-02-28 collective_sun/scripts/operator_product_methods.py

@@ -179,9 +179,9 @@ def symmetrize_operator(oper):
     subsystems = oper.ndim//2
     def _permuted_oper(perm):
         return np.transpose(oper, perm + tuple( np.array(perm) + subsystems ) )
+    perm_sum = sum( _permuted_oper(perm) for perm in it.permutations(range(subsystems)) )
     num_perms = np.math.factorial(subsystems)
-    return sum( _permuted_oper(perm)
-                for perm in it.permutations(range(subsystems)) ) / num_perms
+    return perm_sum / num_perms
 
 # contract operators according to a set diagram
 def contract_ops(base_ops, diagram):
@@ -319,7 +319,7 @@ def evaluate_multi_local_op(local_op, pops_lft, pops_rht = None, diagonal = None
         if any( pop < 0 for pop in remainder ): continue
 
         assignment_rht = _remainder(pops_rht, remainder)
-        if any(np.array(assignment_rht) < 0): continue
+        if any( assignment < 0 for assignment in assignment_rht ): continue
 
         # TODO: compute remainder_perms in a better way
         remainder_perms = multinomial(remainder) / norm

### 2020-02-28 collective_sun/scripts/shell_model.py

@@ -11,7 +11,7 @@ from operator_product_methods import reduced_diagrams, operator_contractions, \
 np.set_printoptions(linewidth = 200)
 cutoff = 1e-10
 
-lattice_shape = (3,3)
+lattice_shape = (10,10)
 alpha = 3 # power-law couplings ~ 1 / r^\alpha
 
 # values of the ZZ coupling to simulate in an XXZ model
@@ -184,12 +184,8 @@ def to_proj(vec):
 
 # decompose couplings in the basis of coefficients that generate eigenstates
 for shell in range(shell_num):
-    sunc[shell] = {}
-    sunc[shell]["disp"] = to_proj( eigs[shell]["disp"] ) @ sunc["disp"]
-    sunc[shell]["mat"] = disp_vec_to_mat(sunc[shell]["disp"])
-    sunc[shell]["pair"] = mat_to_pair_vec(sunc[shell]["mat"])
-    sunc[shell]["col"] = sum(sunc[shell]["mat"])
-    sunc[shell]["tot"] = sum(sunc[shell]["pair"])
+    sunc_shell_disp = to_proj( eigs[shell]["disp"] ) @ sunc["disp"]
+    sunc[shell] = disp_vec_to_mat(sunc_shell_disp)
 
 ##########################################################################################
 # compute operators in the spin projection/shell basis
@@ -219,16 +215,16 @@ Z2_products[2] = np.zeros((spin_num+1,shell_num))
 Z2_products[3] = np.zeros((spin_num+1,shell_num,shell_num))
 
 for shell in range(shell_num):
-    couplings = [ sunc[shell]["mat"], sunc["mat"] ]
+    couplings = [ sunc[shell], sunc["mat"] ]
     Z2_products[2][:,shell] \
         = evaluate_operator_product(couplings, Z2_diags[2], Z2_opers[2])
 
-    couplings = [ sunc[shell]["mat"], sunc["mat"], sunc[shell]["mat"] ]
+    couplings = [ sunc[shell], sunc["mat"], sunc[shell] ]
     Z2_products[3][:,shell,shell] \
         = evaluate_operator_product(couplings, Z2_diags[3], Z2_opers[3])
 
 for shell_lft, shell_rht in itertools.combinations(range(shell_num), 2):
-    couplings = [ sunc[shell_lft]["mat"], sunc["mat"], sunc[shell_rht]["mat"] ]
[…]

### 2020-03-03 collective_sun/sun_perturbation.tex
added discussion motivating the calculation of operator products restricted to the PS manifold

@@
 \newcommand{\M}{\mathcal{M}}
+\newcommand{\N}{\mathcal{N}}
 \renewcommand{\O}{\mathcal{O}}
 \renewcommand{\P}{\mathcal{P}}
+\newcommand{\Q}{\mathcal{Q}}
 \renewcommand{\S}{\mathcal{S}}
@@
   &&
-  \Pi_{pq} \equiv \sum_{\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)},
+  \Pi_{pq} \equiv \sum_{\mu,\nu} s_{\mu\nu}^{(p)} s_{\nu\mu}^{(q)},
   \label{eq:ints}
@@
 single spin; $h_{pq}$ are scalar coefficients; the operator
-$S_{\mu\nu}^{(p)}\equiv\op{\mu}{\nu}_p$ flips the state of spin $p$ to
+$s_{\mu\nu}^{(p)}\equiv\op{\mu}{\nu}_p$ flips the state of spin $p$ to
 $\ket\mu$ from $\ket\nu$; and the operator $\Pi_{pq}$ permutes the
@@
 multilevel generalization of the SU(2)-symmetric spin interaction
-$\v S_p\c\v S_q$ for $\v S=\p{s_\x,s_\y,s_\z}$, with $s_\alpha$ an
-individual spin operator proportional to the Pauli matrix
-$\sigma_\alpha$.
+$\v s_p\c\v s_q$ for $\v s=\p{s_\x,s_\y,s_\z}$, with
+$s_\alpha=\sigma_\alpha/2$ and $\sigma_\alpha$ a Pauli matrix for an
+individual spin.
 
@@
 where $p,q$ index individual spins; $v_{Xp}$ and $w_{Opq}$ are real
-numbers; $X$ is a trace-zero single-spin operator on $\H_n$; $O$ is a
-trace-zero two-spin operator on $\H_n\otimes\H_n$ that obeys pair-wise
-permutational symmetry, i.e.~$O_{pq}=O_{qp}$; $\v v_X$ and $\v X$ are
-$N$-component vectors defined by
+numbers (coefficients); $X$ is a trace-zero single-spin operator on
+$\H_n$; $O$ is a trace-zero two-spin operator on $\H_n\otimes\H_n$
+that obeys pair-wise permutational symmetry,
+i.e.~$O_{pq} = \Pi_{pq} O_{pq} \Pi_{pq} = O_{qp}$; $\v v_X$ and $\v X$
+are $N$-component vectors defined by
 \begin{align}
@@
 throughout these notes, but note that this restriction can be relaxed
-to the assumption that the initial state of the spins is
+to the assumption that the initial state of all spins is
 permutationally symmetric (e.g.~a spin-polarized state).  We will also
@@
 $h_{pq}=-h/\abs{p-q}^\alpha$ on a $D$-dimensional lattice, for
-example, yield a non-vanishing spectral gap in the thermodynamic limit
-when $\alpha\le D$ (see Appendix \ref{sec:gap}).
+example, always yield a non-vanishing spectral gap when $\alpha\le D$
+(see Appendix \ref{sec:gap}).
 
@@
   \E \equiv \sum_{\Delta\ne0} \f{\P_\Delta}{\Delta},
+  \label{eq:H_M}
 \end{align}
 where $\P_\Delta$ is a projector onto the eigenspace of the
-interaction Hamiltonian $H_0$, with interaction energy $\Delta$ above
+interaction Hamiltonian $H_0$ with interaction energy $\Delta$ above
 that of permutationally symmetric manifold $\M_0$; that is, the
-interaction energy of states in the image of $\P_\Delta$ is
+interaction energy of states in the support of $\P_\Delta$ is
 $E_0+\Delta$, where
@@
 of the state $\ket\psi$ or the choice of a trace-zero single-spin
-operators $X$ used to build $\V_1^\Delta$.  Finding operators
+operator $X$ used to build $\V_1^\Delta$.  Finding operators
 $\V_1^\Delta$ that generate eigenvectors of the interaction
@@
 
-If the permutationally symmetric manifold $\M_0$ is gapped away from
-all other states, then then a vector $\V_1^\Delta\ket\psi$ with
-$\Delta=0$ must lie within the permutationally symmetric manifold
-$\M_0$, which implies that the operator $\V_1^\Delta$ preserves the
-permutational symmetry of $\M_0$.  Indeed, a constant vector
-$\v\1\equiv\p{1,1,1,\cdots}$ of ones satisfies the condition in
-\eqref{eq:cond_1} with eigenvalue $0$, which implies that all vectors
-$\v v$ satisfying \eqref{eq:cond_1} with $\Delta\ne0$ must be
-orthogonal to $\v\1$, and therefore mean-zero.
-
-We now return to the task of computing the second-order effective
-Hamiltonian $H_1^{(2)}$.  Any coefficient vector $\v v_X$ can be
-expanded into its projections $\v v_X^\Delta$ onto the eigenspace of
-vectors satisfying \eqref{eq:cond_1} with eigenvalue $\Delta$,
+% If the permutationally symmetric manifold $\M_0$ is gapped away from
+% all other states, then then a vector $\V_1^\Delta\ket\psi$ with
+% $\Delta=0$ must lie within the permutationally symmetric manifold
+% $\M_0$, which implies that the operator $\V_1^\Delta$ preserves the
+% permutational symmetry of $\M_0$.  Indeed, a constant vector
+% $\v\1\equiv\p{1,1,1,\cdots}$ of ones satisfies the condition in
+% \eqref{eq:cond_1} with eigenvalue $0$, which implies that all
+% vectors $\v v$ satisfying \eqref{eq:cond_1} with $\Delta\ne0$ must
+% be orthogonal to $\v\1$, and therefore mean-zero.
+
+% We now return to the task of computing the second-order effective
+% Hamiltonian $H_1^{(2)}$.
+Any coefficient vector $\v v_X$ can be expanded into its projections
+$\v v_X^\Delta$ onto the eigenspace of vectors satisfying
+\eqref{eq:cond_1} with eigenvalue $\Delta$,
 i.e.~$\v v_X = \sum_\Delta \v v_X^\Delta$, which also allows us to
@@
 where the product $\P_0 \p{\V_1^\Delta}^2 \P_0$ is worked out in
-Appendix \ref{sec:PXYP}.  In total, using the fact that
-$\v v_X^\Delta$ is mean-zero for all $\Delta\ne0$, we find
+Appendix \ref{sec:PXYP}.  In total, we find that
 \begin{align}
@@
 where $\EQPS$ denotes equality under a restriction to the
-permutationally symmetric manifold $\M_0$; $\v v_X^\Delta$ is the
+permutationally symmetric manifold $\M_0$; and $\v v_X^\Delta$ is the
 projection of $\v v_X$ onto the $\Delta$-eigenspace of the matrix
@@
 are an average of $w_{Opq}$ and a collective version of the two-spin
-operator $O_{pq}$.  If the two-spin operator $O=X\otimes Y$, then
+operator $O_{pq}$.  If the two-spin operator
+$O=X\otimes Y+Y\otimes X$, then
 \begin{align}
   H_2^{(1)} \EQPS \f12 \sum_{\p{X,Y}} \bar w_{\p{X,Y}}
-  \p{\col{X}\,\col{Y} - \col{XY}}.
+  \p{\col{X}\,\col{Y} + \col{Y}\,\col{X} - \col{XY} - \col{YX}}.
 \end{align}
@@
 permutation of its tensor factors; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is
-a choice of $M$ distinct spins that $O$ acts on; and $\C_N\p{n}$ is a
-choice of $n$ elements from a set of $N$.
+a choice of $M$ distinct spins that $O$ acts on; and $\C_N\p{n}$ is
+the set of all distinct choices of $n$ elements from a set of size
+$N$.
 
@@
 
-We first consider the single-body case and simplify
+As a toy calculation, we first consider the simple case of a
+single-body perturbation and simplify
 \begin{align}
@@
 
-We wish to simplify
+We now consider the full $M$-body product
 \begin{align}
@@
 $\p{N-1}\times\p{N-1}$ matrices.  We then reduce the two-body
-eigenvalue problem even further for the case of an isotropic system,
-to that of diagonalizing a single $\sim\p{N/2^D}\times\p{N/2^D}$
-matrix.  Finally, we provide an analytic solution to the two-body
-problem in the case of a fully collective model with uniform
-couplings.
+eigenvalue problem even further to that of diagonalizing a single
+$\sim\p{N/2^D}\times\p{N/2^D}$ matrix in the case of an isotropic
+system with a translationally invariant (and isotropic) perturbation.
+Finally, we provide an analytic solution to the two-body problem in
+the case of a fully collective model with uniform couplings.
 
@@
 \begin{align}
-  \ket{a} \equiv \f1{\sqrt{N_a}}
+  \ket{a} \equiv \f1{\sqrt{\N_a}}
   \sum_{\substack{d\in\ZZ_L^D\\\abs{d}=a}} \ket{d},
@@
 implicitly understood to mean the smallest Euclidean distance of $d$
-from the origin; and $N_a$ is a normalization factor equal to the
+from the origin; and $\N_a$ is a normalization factor equal to the
 number of displacements $d\in\ZZ_L^D$ with $\abs{d}=a$, so that
@@
   \ut{\v{\m w}}_0
-  \equiv \sum_{a\in\DD_{LD}} \tilde w_a \sqrt{N_a} \ket{a},
+  \equiv \sum_{a\in\DD_{LD}} \tilde w_a \sqrt{\N_a} \ket{a},
 \end{align}
@@
 \begin{align}
-  \tilde h_{ab} \equiv \f1{\sqrt{N_aN_b}}
+  \tilde h_{ab} \equiv \f1{\sqrt{\N_a\N_b}}
   \sum_{\substack{c,d\in\ZZ_L^D\\\abs{c}=a\\\abs{d}=b}} h_{cd}.
@@
 of a fully collective spin model with uniform couplings:
-$h_{pq}=-1+\delta_{pq}$.  In this case, the single-body excitation
-energies $\Delta_k$ are
+$h_{pq}=-1+\delta_{pq}$, where $\delta_{pq}=1$ if $p=q$ and zero
+otherwise.  In this case, the single-body excitation energies
+$\Delta_k$ are
 \begin{align}
@@
 
-\red{TODO: Discuss motivation, why we care about these products
-  (answer: they appear in perturbation theory, as well as the
-  calculation of matrix elements within the outer-most ``shells'' of
-  fixed net spin).  Give an example of the diagram expansions we're
-  after.}
-
-\subsection{The general case}
+Weak $M$-body perturbations $\V_M$ to SU($n$)-symmetric interactions
+can be treated perturbatively, resulting in an effective Hamiltonian
+on a permutationally symmetric manifold $\M_0$ of $N$ spins.
+Computing this effective Hamiltonian generally requires multiplying
+several operators, and projecting the resulting product onto $\M_0$.
+Even if the perturbation $\V_M$ is too strong for rigorous
+perturbative treatment, as long as it is comparable in magnitude to
+the energetic penalty incurred by breaking permutational symmetry, its
+effects can still be accounted for by an effective theory on the joint
+manifold of permutationally symmetric states together with states of
+``weak'' permutational symmetry breaking, e.g.~the manifold of states
+generated by the action of $M$-body perturbations on permutationally
+symmetric states.  The expansion of operators in a basis for this
+joint manifold likewise requires multiplying several operators and
+projecting the resulting product onto $\M_0$.  Specifically, if
+$\V_M=\sum_{k\in\C_N\p{M}}w_kO_k$, then we can define operators
+\begin{align}
+  \V_M^{(k)} \equiv \sum_{p\in\C_N\p{M}} w^{(k)}_p O_p,
+\end{align}
+where $\set{w^{(k)}}$ forms a complete basis for the set of symmetric
+tensors of dimension $M$ with zeros on all diagonals (i.e.~zero in all
+entries with a repeated index).  Denoting a complete basis of
+permutationally symmetric states by $\set{\ket{m}}$, we can then
+identify the $M$-body perturbed states
+\begin{align}
+  \ket{mk} \equiv \f1{\sqrt{\N_{mk}}} \V_M^{(k)} \ket{m},
+  &&
+  \N_{mk} \equiv \bk{m | \V_M^{(k)} \V_M^{(k)} | m}.
+\end{align}
+The restriction of a multi-body operator $\Q=\sum_p v_p Q_p$ to this
+expanded Hilbert space is then
+\begin{align}
+  \Q \simeq_M \P_M \Q \P_M,
+  &&
+  \P_M \equiv \sum_m \op{m} + \sum_{m,k} \op{mk},
+\end{align}
+where finding matrix elements of $\O$ now requires computing products
+of the form
+\begin{align}
+  \bk{\ell | \V_M^{(j)} \Q \V_M^{(k)} | m}
+  = \sum_{p,q,r} w^{(j)}_p v_q w^{(k)}_r \bk{\ell | O_p Q_q O_r | m}.
+  \label{eq:example_prod}
+\end{align}
+The purpose of this section is to work out such products in a
+systematic way.  To this end, we derive a general diagrammatic
+expansion of these products, as well as rules to simplify the
+resulting expansions so as to minimize their computational cost.
+These diagrammatic methods allow us to algorithmically evaluate
+products such as \eqref{eq:example_prod} from the bare multi-body
+operator data $\p{w^{(j)},O},\p{v,Q},\p{w^{(k)},O}$.  We then work out
+some special cases of products that are relevant to the perturbation
+theory in Sections \ref{sec:single_body_pert} and
+\ref{sec:two_body_pert}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{The general case: a diagrammatic expansion}
 
@@
 indices according to choices of $g_S$ for all $S\in\PP\p{\ZZ_p}$ by
-q$g$, where $\PP\p{\ZZ_p}$ is the power set (i.e.~set of all subsets)
+$g$, where $\PP\p{\ZZ_p}$ is the power set (i.e.~set of all subsets)
 of $\ZZ_p$.  Given a vector $\m M$ of the dimensions $M_j$, we then

### 2020-03-05 collective_sun/sun_perturbation.tex
adding section on multi-body eigenvalue problem with translational invariance

@@
 \renewcommand{\S}{\mathcal{S}}
+\newcommand{\T}{\mathcal{T}}
 \newcommand{\V}{\mathcal{V}}
@@
 
-\usepackage{accents}
+\usepackage{accents} % for the undertilde
 \newcommand{\ut}{\undertilde}
@@
 \def\oop#1#2{\oket{#1}\!\obra{#2}}
+
+\usepackage{mathtools} % for coloneqq
 
@@
   \p{\m h - \diag\v h}\c\v v = \Delta \v v,
-  &&
-  \m h \equiv \sum_{p,q} h_{pq} \op{p}{q},
-  &&
-  \v h \equiv \sum_q h_q \ket{q},
   \label{eq:cond_1}
 \end{align}
-where $\m h$ is a matrix of all $h_{pq}$, and
-$\diag\v h\equiv\sum_p h_p \op{p}$ is a matrix with $\v h$ on the
-diagonal and zeros everywhere else.  Solving the general eigenvalue
-problem in \eqref{eq:cond_1} requires diagonalizing the $N\times N$
-matrix $\m h-\diag\v h$.  When the interaction Hamiltonian $H_0$ is
-translationally invariant, however, this eigenvalue problem can be
-solved analytically; we provide the corresponding solution in Appendix
-\ref{sec:trans_inv}.
+where $\m h\equiv\sum_{p,q}h_{pq}\op{p}{q}$ is a matrix of all
+couplings $h_{pq}$; the vector $\v h\equiv\sum_{p,q}h_{pq}\ket{p}$ is
+the sum of all columns of $\m h$; and $\diag\v h$ is a matrix with
+$\v h$ on the diagonal and zeroes everywhere else.  Solving the
+general eigenvalue problem in \eqref{eq:cond_1} requires diagonalizing
+the $N\times N$ matrix $\m h-\diag\v h$.  When the interaction
+Hamiltonian $H_0$ is translationally invariant, however, this
+eigenvalue problem can be solved analytically; we provide the
+corresponding solution in Appendix \ref{sec:trans_inv}.
 
@@
   + \sum_O \sum_{p<q}
-  \sp{\sum_k \p{h_{kp} w_{Okq} w_{Okq} + h_{kq} w_{Opk}}
+  \sp{\sum_k \p{h_{pk} w_{Okq} + h_{qk} w_{Opk}}
   - \bar h_{pq} w_{Opq}}
@@
   \equiv \sum_{p<q} \ket{\set{p,q}} \sp{\sum_{k\notin\set{p,q}}
-    \p{h_{kp} \bra{\set{k,q}} + h_{kq} \bra{\set{p,k}}}
+    \p{h_{pk} \bra{\set{k,q}} + h_{qk} \bra{\set{p,k}}}
     - \bar h_{pq} \bra{\set{p,q}}}.
@@
 \end{align}
-Solving the general eigenvalue problem in \eqref{eq:cond_2} requires
-diagonalizing the ${N \choose 2}\times{N \choose 2}\sim N^2\times N^2$
-matrix $\check{\m h}$.  When both he interaction Hamiltonian $H_0$
+We explicitly exclude $k\in\set{p,q}$ in \eqref{eq:h_super_mat}
+because these terms vanish in \eqref{eq:H_V2_psi} from the fact that
+$h_{kk}=w_{Okk}=0$.  Solving the general eigenvalue problem in
+\eqref{eq:cond_2} requires diagonalizing the
+${N \choose 2}\times{N \choose 2}\sim N^2\times N^2$ matrix
+$\check{\m h}$.  When both he interaction Hamiltonian $H_0$
 translationally invariant, however, this eigenvalue problem can be
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Constructing eigenstates of multi-body perturbations}
+\section{Constructing multi-body excitation energy eigenstates}
 \label{sec:H_VM_psi}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Single-body perturbation}
+\subsection{Single-body excitations}
 \label{sec:H_V1_psi}
@@
 \end{align}
+The action of a single-body perturbation $\V_1=\sum_p v_p X_p$ on a
+fully symmetric state therefore generates an eigenstate of $H_0$ with
+energy $E_0+\Delta$ if the vector $\v v\equiv\sum_p v_p\ket{p}$
+satisfies the eigenvalue equation
+\begin{align}
+  \p{\m h - \diag\v h} \c \v v = \Delta \v v,
+  \label{eq:H_V1_psi_eig}
+\end{align}
+where $\m h\equiv\sum_{p,q}h_{pq}\op{p}{q}$ is a matrix of all
+couplings $h_{pq}$; the vector $\v h\equiv\sum_{p,q}h_{pq}\ket{p}$ is
+the sum of all columns of $\m h$; and $\diag\v h$ is a matrix with
+$\v h$ on the diagonal and zeroes everywhere else.
+
+% When the interaction Hamiltonian $H_0$ is translationally invariant,
+% the single-body eigenvalue problem in \eqref{eq:H_V1_psi_eig} is
+% exactly solvable.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Multi-body perturbation}
+\subsection{Multi-body excitations}
 \label{sec:H_VM_psi}
@@
   &&
-  h_p \equiv \sum_q h_{pq}.
-\end{align}
-The terms in \eqref{eq:H_VM_psi_start} with only one of $p\in k$ or
-$q\in k$, meanwhile, are
+  h_p \equiv \sum_q h_{pq},
+\end{align}
+which implies that the terms in \eqref{eq:H_VM_psi_start} with
+$p,q\in k$ or $p,q\notin k$ are
+\begin{align}
+  \sum_{k\in\C_N\p{M}} \p{E_0 - \bar h_k} w_k O_k \ket\psi
+  = E_0 \V_M \ket\psi
+  - \sum_{k\in\C_N\p{M}} \p{\bar{\m h} * \m w}_k O_k \ket\psi,
+\end{align}
+where $\bar{\m h}$ is a tensor with entries $\bar h_k$; and
+$\m x*\m y$ denotes an element-wise (Kronecker) product of $\m x$ and
+$\m y$, with entries $x_ky_k$.
+
+The remaining terms in \eqref{eq:H_VM_psi_start} with only one of
+$p\in k$ or $q\in k$ are
 \begin{align}
@@
 \end{align}
-Defining
-\begin{align}
-  \p{h\circ w}_k
+For each fixed $j$, the sum over $p$ above is essentially a
+contraction of the matrix $\m h$ with the $j$-th index of the tensor
+$\m w$.  Denoting such a contraction by $\m h\c_j\m w$, we can
+therefore write this result as
+\begin{align}
+  \sum_{k\in\C_N\p{M}} \p{\m h\circ \m w}_k O_k \ket\psi,
+  &&
+  \m h\circ\m w \equiv \sum_{j=1}^M \m h\c_j\m w,
+\end{align}
+where, for clarity, the matrix elements of $\m h\circ\m w$ are
+\begin{align}
+  \p{\m h\circ\m w}_k
   \equiv \sum_{\substack{p\in\ZZ_N\\p\notin k}} \sum_{j=1}^M
@@
     + h_{k_2 p} w_{k_1pk_3\cdots k_M}
-    + \cdots + h_{k_M p} w_{k_1k_2\cdots k_{M-1} p}},
-\end{align}
-we thus find that
+    + \cdots + h_{k_M p} w_{k_1k_2\cdots k_{M-1} p}}.
+\end{align}
+Altogether, we thus find that
 \begin{align}
@@
   = E_0 \V_M \ket\psi + \sum_{k\in\C_N\p{M}}
-  \sp{\p{h\circ w}_k - \bar h_k w_k} O_k \ket\psi.
-\end{align}
-If $M=1$, then $\p{h\circ w}_k=\sum_\ell h_{k\ell}w_\ell$ and
+  \p{\m h\circ\m w - \bar{\m h} * \m w}_k O_k \ket\psi.
+  \label{eq:H_VM_psi_end}
+\end{align}
+The action of the $M$-body perturbation $\V_M$ on a fully symmetric
+state therefore generates an eigenstate of $H_0$ with energy
+$E_0+\Delta$ if the coefficient tensor $\m w$ satisfies the
+generalized eigenvalue equation
+\begin{align}
+  \m h\circ\m w - \bar{\m h} * \m w = \Delta \m w.
+  \label{eq:H_VM_psi_eig}
+\end{align}
+If $M=1$, then $\p{\m h\circ\m w}_k=\sum_\ell h_{k\ell}w_\ell$ and
 $\bar h_k=h_k$, so
@@
 which is precisely what we found in \eqref{eq:H_V1_psi_end}.  If
-$M=2$, then $\p{h\circ w}_{pq}=\sum_k\p{h_{kp} w_{kq}+h_{kq}w_{pk}}$,
-so
+$M=2$, then
+$\p{\m h\circ\m w}_{pq}=\sum_k\p{h_{pk}w_{kq}+h_{qk}w_{pk}}$, so
 \begin{align}
@@
   = E_0 \V_2 \ket\psi + \sum_{\p{p,q}\in\C_N\p{2}}
-  \sp{\sum_{k\in\ZZ_N}\p{h_{kp} w_{kq} + h_{kq} w_{pk}}
+  \sp{\sum_{k\in\ZZ_N}\p{h_{pk} w_{kq} + h_{qk} w_{pk}}
     - \bar h_{pq} w_{pq}}
@@
 where $\bar h_{pq}=h_p+h_q-2h_{pq}$ with $h_p\equiv\sum_qh_{pq}$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Translational invariance}
+\label{sec:H_VM_psi_TI}
+
+Solving eigenvalue problem in \eqref{eq:H_VM_psi_eig} to find $M$-body
+perturbations $\V_M$ that generate states of definite interaction
+energy nominally requires diagonalizing an
+${N\choose M}\times{N\choose M}\sim N^M\times N^M$ matrix.  Here, we
+simplify this problem to that of diagonalizing an
+$\sim{N-1\choose M-1}\times{N-1\choose M-1}\sim N^{M-1}\times N^{M-1}$
+matrix when both the interaction Hamiltonian $H_0$ and the
+perturbation $\V_M$ are translationally invariant.  This
+simplification reduces the computational cost of solving the $M$-body
+eigenvalue problem by a factor of $\O\p{N^3}$.
+
+Our task is essentially to express the multi-body eigenvalue problem
+in \eqref{eq:H_VM_psi_eig} in a basis that obeys both permutational
+symmetry and translational invariance.  To this end, we first expand
+the coupling tensor $\m w$ in a complete basis for all tensors of
+dimension $M$:
+\begin{align}
+  \m w = \sum_{k\in\ZZ_N^M} w_k \m e_k,
+\end{align}
+where $\m e_k$ is a unit tensor with one in the entry indexed by $k$
+and zeros everywhere else.  We then use the permutational symmetry of
+$\m w$ to expand it in a basis for symmetric tensors, making use of
+the fact that all entries of $\m w$ with repeated indices are zero:
+\begin{align}
+  \m w = \sum_{k\in\C_N\p{M}} w_k \m m_k,
+  &&
+  \m m_k \equiv \sum_{\sigma\in P_M} \m e_{\sigma\p{k}},
+  \label{eq:sym_tensor}
+\end{align}
+where $P_M$ is the set of all permutations of $M$ symbols.  On a
+$D$-dimensional lattice of linear size $L$, translational invariance
+implies that the coefficients $w_k$ remain the same if all spins
+indexed by $k$ are shifted by the same displacement $d\in\ZZ_L^D$.  To
+account for this invariance, we denote a translation of all spins in
+$k$ by a displacement $d\in\ZZ_L^D$ by $k_d$, and define the
+translation-symmetrized basis tensors
+\begin{align}
+  \m m_k^{\t{TS}}
+  \equiv \f1{\N_k^{\t{TS}}} \sum_{d\in\ZZ_L^D} \m m_{k_d}
+  = \f1{\N_k^{\t{TS}}} \sum_{d\in\ZZ_L^D} \sum_{\sigma\in P_M}
+  \m e_{\sigma(k_d)},
+\end{align}
+where the normalization factor $\N_k^{\t{TS}}$ enforces that the
+entries of $\m m_k^{\t{TS}}$ are all either zero or one, which
+otherwise might not occur if some translations of $k$ results in the
+same choice of spins.  We then define the set $\T_N\p{M}$ of
+equivalence classes in $\C_N\p{M}$ under arbitrary translations, and
+expand
+\begin{align}
+  \m w = \sum_{k\in\T_N\p{M}} w_k \m m_k^{\t{TS}}.
+  \label{eq:sym_shift_tensor}
+\end{align}
+Having expanded $\m w$ in a basis that respects permutational symmetry
+and translational invariance, we need to express the eigenvalue
+equation \eqref{eq:H_VM_psi_eig} in the same basis.  In order to do
+so, we label spins $p,q$ by the position of the corresponding lattice
+site in $\ZZ_L^D$, and use translational invariance of the couplings
+$\m h$ to expand
+\begin{align}
+  \m h \equiv \sum_{p,q\in\ZZ_L^D} h_{pq} \op{p}{q}
+  = \sum_{q,d\in\ZZ_L^D} h_{q+d,q} \op{q+d}{q}
+  = \sum_{d\in\ZZ_L^D} h_{d,0} \m s_d,
+  &&
+  \m s_d \equiv \sum_{q\in\ZZ_L^D} \op{q+d}{d}.
+\end{align}
+Starting from this form of $\m h$ and the expansion of $\m w$ in
+\eqref{eq:sym_tensor}, we can work out that
+\begin{align}
+  \m h\circ\m w
+  = \sum_{k\in\T_N\p{M}} \m m_k^{\t{TS}}
+  \sum_{d\in\ZZ_L^D} h_{d,0} \sum_{a=1}^M w_{k_{d;a}},
+\end{align}
+where $k_{d;a}$ denotes the spins acquired by displacing the $a$-th
+spin of $k$ by $d$.  With a bit of additional work, we can also
+simplify
+\begin{align}
+  \bar{\m h} * \m w = M h_0 \m w
+  - \sum_{k\in\T_N\p{M}} w_k \m m_k^{\t{TS}}
+  \sum_{p,q\in k} h_{pq}
+  &&
+  h_0 \equiv \sum_{d\in\ZZ_L^D} h_{d,0}.
+\end{align}
+Putting everything together, the generalized eigenvalue problem in
+\eqref{eq:H_VM_psi_eig} can be written in the form
+\begin{align}
+  \sum_{k\in\T_N\p{M}} \m m_k^{\t{TS}}
+  \sp{\sum_{p,q\in k} h_{pq} w_k
+    + \sum_{d\in\ZZ_L^D} h_{d,0} \sum_{a=1}^M w_{k_{d;a}}}
+  = \p{M h_0 + \Delta} \m w,
+\end{align}
+or equivalently
+\begin{align}
+  \ut{\check{\m h}} \c \v{\m w} = \p{M h_0 + \Delta} \v{\m w},
+  \label{eq:H_VM_eig_simp}
+\end{align}
+where
+\begin{align}
+  \ut{\check{\m h}} \equiv
+  \sum_{k\in\T_N\p{M}} \ket{k}
+  \sp{\sum_{p,q\in k} h_{pq} \bra{k}
+    + \sum_{d\in\ZZ_L^D} h_{d,0} \sum_{a=1}^M \bra{k_{d;a}}},
+  &&
+  \v{\m w} \equiv \sum_{k\in\T_N\p{M}} w_k \ket{k},
+\end{align}
+and, to keep notation simple, we define $\ket{k}=0$ for all $k$ with
+repeated values, and enforce $\ket{j}=\ket{k}$ if spins $j$ and spins
+$k$ are the same up to a global translation.
+
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \subsubsection{Two-body excitation energy eigenstates}
+% \label{sec:trans_inv_two}
+
+% The single-body excitation energy eigenvalue problem in
+% \eqref{eq:H_V1_psi_eig} is solvable analytically when the interaction
+% Hamiltonian $H_0$ is translationally invariant.
+
+\vspace{1cm}
+
+\red{TODO: write codes to check above work, fix any errors, explicitly
+  write out single- and two-body reduced eigenvalue problems, include
+  sub-section about isotropy, replace the following appendix}
+
+\vspace{1cm}
 
@@
 \end{align}
-where finding matrix elements of $\O$ now requires computing products
+where finding matrix elements of $\Q$ now requires computing products
 of the form
@@
 \begin{align}
-  \S\p{g} \equiv \sp{\prod_{j\in\ZZ_p} \C\p{ g_S : j\in S }}
+  \S\p{g} \equiv \sp{\prod_{j\in\ZZ_p} { M_j \choose g_S : j\in S }}
   \sp{\prod_{S\in\PP\p{\ZZ_p}} \p{g_S!}^\abs{S}}
@@
 \begin{align}
-  \C\p{a_1,a_2,\cdots,a_m}
-  \equiv \f{\p{a_1+a_2+\cdots+a_m}!}{a_1!a_2!\cdots a_m!}
+  {M_j \choose a_1,a_2,\cdots,a_m}
+  \equiv \f{M_j!}{a_1!a_2!\cdots a_m!}
 \end{align}
 that account for the number of ways to partition the indices of each
-$w_j,O_j$ into sets of shared indices as specified by the values of
-$g_S$ for all $j\in S$, and the second product in
-\eqref{eq:sym_prod_symmetry} accounts for the number of ways to
-permute each set of shared indices on every tensor.  Altogether, the
-symmetry factors $\S\p{g}$ cancel out with the $\sim M_j!$ prefactors
-in \eqref{eq:sym_prod_vec}, so
+$w_j,O_j$ into sets of shared indices as specified by $g$, and the
+second product in \eqref{eq:sym_prod_symmetry} accounts for the number
+of ways to permute each set of shared indices on every tensor.
+Altogether, the symmetry factors $\S\p{g}$ cancel out with the
+$\sim M_j!$ prefactors in \eqref{eq:sym_prod_vec}, so
 \begin{align}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Matrix elements of multi-body operators}
+
+\red{TODO: write section}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Two single-body operators}
@@
 $v_{k\ell},w_{pq}$ are scalar coefficients with $m_{pq}=m_{qp}$ and
-$m_{pp}=0$ for each of $m\in\set{v,w}$; and $O,Q$ are
-permutationally-symmetric two-spin operators.  This product appears,
-for example, in the calculation of the second-order effective
-Hamiltonian $H_2^{(2)}$ induced on the permutationally symmetric
-manifold $\M_0$ by the two-body perturbation $\V_2$ in
-\eqref{eq:perturbations} in \eqref{eq:perturbations}.  Defining
+$m_{pp}=0$ for each of $m\in\set{v,w}$; and $O,Q$ are permutationally
+symmetric two-spin operators.  This product appears, for example, in
+the calculation of the second-order effective Hamiltonian $H_2^{(2)}$
+induced on the permutationally symmetric manifold $\M_0$ by the
+two-body perturbation $\V_2$ in \eqref{eq:perturbations} in
+\eqref{eq:perturbations}.  Defining
 \begin{align}

### 2020-03-05 collective_sun/sun_perturbation.tex
removed obsolete sections and fixed references appropriately

@@
   \Pi_{pq} \equiv \sum_{\mu,\nu} s_{\mu\nu}^{(p)} s_{\nu\mu}^{(q)},
-  \label{eq:ints}
+  \label{eq:H_0}
 \end{align}
@@
 example, always yield a non-vanishing spectral gap when $\alpha\le D$
-(see Appendix \ref{sec:gap}).
+(see Appendix \ref{sec:spectral_gap}).
 
@@
   \p{\m h - \diag\v h}\c\v v = \Delta \v v,
-  \label{eq:cond_1}
+  \label{eq:V1_eig}
 \end{align}
@@
 $\v h$ on the diagonal and zeroes everywhere else.  Solving the
-general eigenvalue problem in \eqref{eq:cond_1} requires diagonalizing
+general eigenvalue problem in \eqref{eq:V1_eig} requires diagonalizing
 the $N\times N$ matrix $\m h-\diag\v h$.  When the interaction
@@
 eigenvalue problem can be solved analytically; we provide the
-corresponding solution in Appendix \ref{sec:trans_inv}.
+corresponding solution at the end of Appendix \ref{sec:H_V1_psi}.
 
 If we {\it construct} a perturbation $\V_1^\Delta$ with coefficients
-$\v v_X^\Delta$ that satisfy \eqref{eq:cond_1} with eigenvalue
+$\v v_X^\Delta$ that satisfy \eqref{eq:V1_eig} with eigenvalue
 $\Delta$, then
@@
 
-% If the permutationally symmetric manifold $\M_0$ is gapped away from
-% all other states, then then a vector $\V_1^\Delta\ket\psi$ with
-% $\Delta=0$ must lie within the permutationally symmetric manifold
-% $\M_0$, which implies that the operator $\V_1^\Delta$ preserves the
-% permutational symmetry of $\M_0$.  Indeed, a constant vector
-% $\v\1\equiv\p{1,1,1,\cdots}$ of ones satisfies the condition in
-% \eqref{eq:cond_1} with eigenvalue $0$, which implies that all
-% vectors $\v v$ satisfying \eqref{eq:cond_1} with $\Delta\ne0$ must
-% be orthogonal to $\v\1$, and therefore mean-zero.
-
-% We now return to the task of computing the second-order effective
-% Hamiltonian $H_1^{(2)}$.
 Any coefficient vector $\v v_X$ can be expanded into its projections
 $\v v_X^\Delta$ onto the eigenspace of vectors satisfying
-\eqref{eq:cond_1} with eigenvalue $\Delta$,
+\eqref{eq:V1_eig} with eigenvalue $\Delta$,
 i.e.~$\v v_X = \sum_\Delta \v v_X^\Delta$, which also allows us to
@@
 $H_2^{(2)}$, as before we pick an arbitrary permutationally symmetric
-state $\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_VM_psi})
+state $\ket\psi\in\M_0$ and expand (see Appendix
+\ref{sec:eigenstates})
 \begin{align}
@@
   \check{\m h} \c \v{\m w} = \Delta \v{\m w},
-  \label{eq:cond_2}
+  \label{eq:V2_eig}
 \end{align}
@@
     - \bar h_{pq} \bra{\set{p,q}}}.
-  \label{eq:h_super_mat}
-\end{align}
-We explicitly exclude $k\in\set{p,q}$ in \eqref{eq:h_super_mat}
-because these terms vanish in \eqref{eq:H_V2_psi} from the fact that
+  \label{eq:V2_eig_mat}
+\end{align}
+We explicitly exclude $k\in\set{p,q}$ in \eqref{eq:V2_eig_mat} because
+these terms vanish in \eqref{eq:H_V2_psi} from the fact that
 $h_{kk}=w_{Okk}=0$.  Solving the general eigenvalue problem in
-\eqref{eq:cond_2} requires diagonalizing the
+\eqref{eq:V2_eig} requires diagonalizing the
 ${N \choose 2}\times{N \choose 2}\sim N^2\times N^2$ matrix
@@
 reduced to that of diagonalizing an $\sim N\times N$ matrix; we
-discuss this reduction in Appendix \ref{sec:trans_inv}.
+perform this reduction in Appendix \ref{sec:eigenstates_TI}.
 
 If we {\it construct} a perturbation $\V_2^\Delta$ with coefficients
-$\v{\m w}_O^\Delta$ that satisfy \eqref{eq:cond_2} with eigenvalue
+$\v{\m w}_O^\Delta$ that satisfy \eqref{eq:V2_eig} with eigenvalue
 $\Delta$, then
@@
 $\v{\m w}_O$ into its projections $\v{\m w}_O^\Delta$ onto the
-eigenspace of vectors satisfying \eqref{eq:cond_2} with eigenvalue
+eigenspace of vectors satisfying \eqref{eq:V2_eig} with eigenvalue
 $\Delta$, i.e.~$\v{\m w}_O=\sum_\Delta\v{\m w}_O^\Delta$, which allows
@@
 of $\v{\m w}$ onto the $\Delta$-eigenspace of the matrix
-$\check{\m h}$ defined in \eqref{eq:h_super_mat} by $\v{\m w}_\Delta$,
+$\check{\m h}$ defined in \eqref{eq:V2_eig_mat} by $\v{\m w}_\Delta$,
 and define
@@
 \section{Existence of a spectral gap with power-law interactions}
-\label{sec:gap}
+\label{sec:spectral_gap}
 
 Here we show that power-law SU($n$)-symmetric interactions of the form
-in \eqref{eq:ints} with $h_{pq}=-h/\abs{p-q}^\alpha$ yield a
+in \eqref{eq:H_0} with $h_{pq}=-h/\abs{p-q}^\alpha$ yield a
 non-vanishing many-body interaction energy gap when $\alpha\le D$,
@@
 $N=L^D$ spins, the spectral gap of the interaction Hamiltonian is (see
-Appendix \ref{sec:trans_inv_single})
+Appendix \ref{sec:H_V1_psi})
 \begin{align}
@@
 \section{Constructing multi-body excitation energy eigenstates}
-\label{sec:H_VM_psi}
+\label{sec:eigenstates}
 
@@
 multi-body perturbation $\V_M$ generates eigenstates of the
-SU($n$)-symmetric interaction Hamiltonian $H_0$ in \eqref{eq:ints}
-when acting on a permutationally symmetric state $\ket\psi\in\M_0$.
-Here $\V_M$ is an $M$-body perturbation of the form
+SU($n$)-symmetric interaction Hamiltonian $H_0$ in \eqref{eq:H_0} when
+acting on a permutationally symmetric state $\ket\psi\in\M_0$.  Here
+$\V_M$ is an $M$-body perturbation of the form
 \begin{align}
@@
 
-% When the interaction Hamiltonian $H_0$ is translationally invariant,
-% the single-body eigenvalue problem in \eqref{eq:H_V1_psi_eig} is
-% exactly solvable.
+When the interaction Hamiltonian $H_0$ is translationally invariant,
+the single-body eigenvalue problem in \eqref{eq:H_V1_psi_eig} is
+exactly solvable.  In this case, the couplings $h_{pq}$ depend only on
+the separation $\abs{p-q}$, so the eigenvectors of $\m h$ are plane
+waves of the form
+\begin{align}
+  \v v_k \equiv \sum_p e^{ip\c k} \ket{p},
+\end{align}
+where on a $D$-dimensional periodic lattice of $N=L^D$ spins, lattice
+sites are indexed by vectors
+$p\in\ZZ_L^D\simeq\set{0,1,\cdots,L-1}^{\otimes D}$, and wavenumbers
+take on values $k\in\ZZ_L^D\times2\pi/L$.  The corresponding
+eigenvalues of $\m h$ can be determined by expanding
+\begin{align}
+  \m h\c\v v_k = \sum_{p,q} h_{pq} e^{iq\c k} \ket{p}
+  = \sum_{p,d} h_{p,p+d} e^{i\p{p+d}\c k} \ket{p}
+  = \sum_d h_{0,d} e^{id\c k} \v v_k
+  = \sum_d h_{0,d} \cos\p{d\c k} \v v_k,
+\end{align}
+where the imaginary terms vanish because $h_{0,d}=h_{0,-d}$.  The
+remainder of \eqref{eq:H_V1_psi_eig} that we need to sort out is
+$\diag\v h$, where all $h_p=\sum_qh_{pq}=\sum_dh_{0,d}$ are equal,
+which implies that $\diag\v h=\sum_dh_{0,d}$ is a scalar.  We thus
+find that,
+\begin{align}
+  \p{\m h - \diag\v h}\c\v v_k = \Delta_k \v v_k,
+  &&
+  \Delta_k \equiv \sum_d h_{0,d} \sp{\cos\p{d\c k}-1}.
+\end{align}
 
@@
 \subsection{Multi-body excitations}
-\label{sec:H_VM_psi}
+\label{sec:eigenstates}
 
@@
 \subsection{Translational invariance}
-\label{sec:H_VM_psi_TI}
-
-Solving eigenvalue problem in \eqref{eq:H_VM_psi_eig} to find $M$-body
-perturbations $\V_M$ that generate states of definite interaction
-energy nominally requires diagonalizing an
+\label{sec:eigenstates_TI}
+
+Solving the eigenvalue problem in \eqref{eq:H_VM_psi_eig} to find
+$M$-body perturbations $\V_M$ that generate states of definite
+interaction energy nominally requires diagonalizing an
 ${N\choose M}\times{N\choose M}\sim N^M\times N^M$ matrix.  Here, we
@@
 
-% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \subsubsection{Two-body excitation energy eigenstates}
-% \label{sec:trans_inv_two}
-
-% The single-body excitation energy eigenvalue problem in
-% \eqref{eq:H_V1_psi_eig} is solvable analytically when the interaction
-% Hamiltonian $H_0$ is translationally invariant.
-
 \vspace{1cm}
@@
 \red{TODO: write codes to check above work, fix any errors, explicitly
-  write out single- and two-body reduced eigenvalue problems, include
-  sub-section about isotropy, replace the following appendix}
+  write out two-body reduced eigenvalue problems, include sub-section
+  about isotropy, replace the following appendix}
 
 \vspace{1cm}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Constructing eigenstates of translationally invariant
-  systems}
-\label{sec:trans_inv}
-
-Here we simplify the eigenvalue problems that appear at second order
-in the single- and two-body theory in Sections
-\ref{sec:single_body_pert} and \ref{sec:two_body_pert} in the case of
-a translationally invariant system with periodic boundary conditions.
-We solve the single-body eigenvalue problem analytically, and reduce
-the two-body eigenvalue problem to that of diagonalizing
-$\p{N-1}\times\p{N-1}$ matrices.  We then reduce the two-body
-eigenvalue problem even further to that of diagonalizing a single
-$\sim\p{N/2^D}\times\p{N/2^D}$ matrix in the case of an isotropic
-system with a translationally invariant (and isotropic) perturbation.
-Finally, we provide an analytic solution to the two-body problem in
-the case of a fully collective model with uniform couplings.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Single-body eigenvalue problem}
-\label{sec:trans_inv_single}
-
-For reference, the single-body eigenvalue problem is
-\begin{align}
-  \p{\m h - \diag\v h}\c\v v = \Delta \v v,
-  \label{eq:cond_1_ref}
-\end{align}
-where $\m h$ is a matrix of all couplings $h_{pq}$ that appear in the
-interaction Hamiltonian $H_0$ in \eqref{eq:ints}; $\v h$ is a vector
-of all $h_p\equiv\sum_q h_{pq}$; $\diag\v h$ is a matrix with $\v h$
-on the diagonal and zeroes everywhere else; and $\v v$ is a vector of
-coefficients used to construct the single-body perturbation
-$\V_1\p{\v v}$ defined in \eqref{eq:perturbations}.  When the
-interaction Hamiltonian $H_0$ is translationally invariant, the
-couplings $h_{pq}$ depend only on the separation $\pm\p{p-q}$, so the
-eigenvectors of $\m h$ are plane waves of the form
-\begin{align}
-  \v v_k \equiv \sum_p e^{ip\c k} \ket{p},
-\end{align}
-where on a $D$-dimensional periodic lattice of $N=L^D$ spins, lattice
-sites are indexed by vectors
-$p\in\ZZ_L^D\simeq\set{0,1,\cdots,L-1}^{\otimes D}$, and wavenumbers
-take on values $k\in\ZZ_L^D\times2\pi/L$.  The corresponding
-eigenvalues of $\m h$ can be determined by expanding
-\begin{align}
-  \m h\c\v v_k = \sum_{p,q} h_{pq} e^{iq\c k} \ket{p}
-  = \sum_{p,d} h_{p,p+d} e^{i\p{p+d}\c k} \ket{p}
-  = \sum_d h_{0,d} e^{id\c k} \v v_k
-  = \sum_d h_{0,d} \cos\p{d\c k} \v v_k,
-\end{align}
-where the imaginary terms vanish because $h_{0,d}=h_{0,-d}$.  The
-remainder of \eqref{eq:cond_1_ref} that we need to sort out is
-$\diag\v h$, where all $h_p=\sum_qh_{pq}=\sum_dh_{0,d}$ are equal,
-which implies that $\diag\v h=\sum_dh_{0,d}$ is a scalar.  We thus
-find that
-\begin{align}
-  \p{\m h - \diag\v h}\c\v v_k = \Delta_k \v v_k,
-  &&
-  \Delta_k \equiv \sum_d h_{0,d} \sp{\cos\p{d\c k}-1}.
-\end{align}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Two-body eigenvalue problem}
-\label{sec:trans_inv_two}
-
-The two-body eigenvalue equation in \eqref{eq:cond_2} is designed to
-find solutions $\m w$ to the equation
-\begin{align}
-  \sum_k \p{h_{pk} w_{kq} + w_{pk} h_{kq}} - \bar h_{pq} w_{pq}
-  = \Delta w_{pq},
-  &&
-  \bar h_{pq} \equiv h_p + h_q - 2h_{pq},
-  &&
-  h_p \equiv \sum_q h_{pq},
-  \label{eq:cond_2_ref_start}
-\end{align}
-where translational invariance implies that
-$\bar h_{pq}=2h_0-2h_{pq}$, so we can equivalently write
-\begin{align}
-  h_{pq} w_{pq}
-  + \f12 \sum_k \p{h_{pk} w_{kq} + w_{pk} h_{kq}}
-  = \p{\f12 \Delta + h_0} w_{pq},
-\end{align}
-which can also be written in matrix form as
-\begin{align}
-  \m h * \m w + \f12 \P_{\t{off-diag}}\p{\m h\c\m w + \m w\c\m h}
-  = \p{\f12 \Delta + h_0} \m w,
-  \label{eq:cond_2_ref}
-\end{align}
-where $\m h$ and $\m w$ are matrices of all $h_{pq}$ and $w_{pq}$;
-$\m h*\m w$ denotes an element-wise (Kronecker) product of $\m h$ and
-$\m w$, i.e.~with matrix elements $h_{pq}w_{pq}$; and
-$\P_{\t{off-diag}}\p{\m m}\equiv \m m-\sum_p\op{p}\m m\op{p}$ denotes
-the projection of $\m m$ onto the space of strictly off-diagonal
-matrices.  In order to find solutions to \eqref{eq:cond_2_ref} when
-the interaction Hamiltonian $H_0$ is translationally invariant, in
-which case $h_{pq}$ depends only on the separation $\pm\p{p-q}$, we
-expand all matrices in a basis that the respects translational
-symmetry.  To this end, for all lattice displacements $d\in\ZZ_L^D$
-and wavenumbers $k\in\ZZ_L^D\times2\pi/L$ we define the {\it cycle
-  matrices}
-\begin{align}
-  \m m_{dk} \equiv \sum_p e^{ip\c k} \op{p+d}{p},
-  \label{eq:cycle}
-\end{align}
-where the element-wise sums in $p+d$ are implicitly taken modulo $L$
-(i.e.~the length of the periodic lattice along each axis).  In terms
-of these cycle matrices, we define
-\begin{align}
-  \tilde w_{dk} \equiv \f1N \tr\p{\m m_{dk}^\dag \m w},
-\end{align}
-and expand
-\begin{align}
-  \m w = \sum_{d,k} \tilde w_{dk} \m m_{dk},
-  &&
-  \m h = \sum_{q,d} h_{q+d,q} \op{q+d}{q}
-  = \sum_d h_{d,0} \m m_{d,0}.
-\end{align}
-We can then write the condition in \eqref{eq:cond_2_ref} as a new set
-of conditions
-\begin{align}
-  \sum_{d\ne0} h_{d,0} \tilde w_{dk} \m m_{dk}
-  + \f12 \sum_{\substack{c,d\\c+d\ne0}}
-  h_{c,0} \tilde w_{dk} \p{1 + e^{-ic\c k}} \m m_{c+d,k}
-  = \p{\f12\Delta + h_0} \sum_{d\ne0} \tilde w_{dk} \m m_{dk},
-\end{align}
-that must hold for all wavenumbers $k$.  With some re-indexing, we can
-write
-\begin{align}
-  \sum_{d\ne0} h_{d,0} \tilde w_{dk} \m m_{dk}
-  + \f12 \sum_{c,d\ne0} h_{d-c,0} \tilde w_{ck}
-  \p{1 + e^{-i\p{d-c}\c k}} \m m_{dk}
-  = \p{\f12\Delta + h_0} \sum_{d\ne0} \tilde w_{dk} \m m_{dk},
-  \label{eq:cond_2_cycle}
-\end{align}
-and define
-\begin{align}
-  \ut{\check{\m h}}_k
-  \equiv \sum_{d\ne0} h_{d,0} \op{d}
-  + \f12 \sum_{c,d\ne0} h_{d-c,0} \p{1+e^{-i\p{d-c}\c k}} \op{d}{c},
-  &&
-  \ut{\v{\m w}}_k \equiv \sum_{d\ne0} \tilde w_{dk} \ket{d},
-  \label{eq:cycle_vecs}
-\end{align}
-in order to express \eqref{eq:cond_2_cycle} as the set of eigenvalue
-equations
-\begin{align}
-  \ut{\check{\m h}}_k \c \ut{\v{\m w}}_k
-  = \p{\f12\Delta + h_0} \ut{\v{\m w}}_k.
-  \label{eq:cond_2_cycle_block}
-\end{align}
-The appearance of a wavenumber index $k$ in
-\eqref{eq:cond_2_cycle_block} implies that all fixed-$k$ components of
-a general coefficient matrix $\m w$ must simultaneously satisfy
-\eqref{eq:cond_2_cycle_block} in order to generate states of fixed
-excitation energy $\Delta$ when the corresponding perturbation
-$\V_2^\Delta$ acts on a permutationally symmetric state
-$\ket\psi\in\M_0$.  The decomposition of an arbitrary coefficient
-vector $\v{\m w}=\sum_\Delta\v{\m w}_\Delta$ into coefficient vectors
-$\v{\m w}_\Delta$ that generate a fixed-excitation-energy perturbation
-$\V_2^\Delta$ is then performed by projecting each $k$-component of
-$\v{\m w}$ onto the corresponding $\Delta$-eigenspace of
-$\ut{\check{\m h}}_k$.  If the perturbation $\V_2$ is translationally
-invariant, then $\ut{\v{\m w}}_k=0$ for all $k\ne0$, so we only need
-to consider the case of $k=0$:
-\begin{align}
-  \ut{\check{\m h}}_0 \c \ut{\v{\m w}}_0
-  = \p{\f12\Delta + h_0} \ut{\v{\m w}}_0,
-  &&
-  \ut{\check{\m h}}_0
-  = \sum_{d\ne0} h_{d,0} \op{d}
-  + \sum_{c,d\ne0} h_{cd} \op{c}{d},
-  &&
-  \ut{\v{\m w}}_0 \equiv \sum_{d\ne0} \tilde w_{d,0} \ket{d}.
-  \label{eq:cond_2_cycle_0}
-\end{align}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Isotropic systems}
-
-If the interaction Hamiltonian $H_0$ and two-body perturbation $\V_2$
-are both translationally invariant isotropic, then we can reduce the
-eigenvalue problem \eqref{eq:cond_2_cycle_0} even further by making
-use of the fact that the couplings $h_{pq}$ depend only on the
-distance $\abs{p-q}$.  To this end, we define uniform superpositions
-of all displacements $d\in\ZZ_L^D$ that are a distance $a\in\RR$ from
-the origin:
-\begin{align}
-  \ket{a} \equiv \f1{\sqrt{\N_a}}
-  \sum_{\substack{d\in\ZZ_L^D\\\abs{d}=a}} \ket{d},
-\end{align}
-where the norm $\abs{d}$ of a vector $d$ on a periodic lattice is
-implicitly understood to mean the smallest Euclidean distance of $d$
-from the origin; and $\N_a$ is a normalization factor equal to the
-number of displacements $d\in\ZZ_L^D$ with $\abs{d}=a$, so that
-$\bk{a|a}=1$.  Defining the set of distances
-\begin{align}
-  \DD_{LD} \equiv \set{ \abs{d} : d \in \ZZ_L^D } \setminus \set{0},
-\end{align}
-we can expand
-\begin{align}
-  \ut{\check{\m h}}_0
-  = \sum_{a\in\DD_{LD}} \tilde h_a \op{a}
-  + \sum_{a,b\in\DD_{LD}} \tilde h_{ab} \op{a}{b},
-  &&
-  \ut{\v{\m w}}_0
-  \equiv \sum_{a\in\DD_{LD}} \tilde w_a \sqrt{\N_a} \ket{a},
-\end{align}
-where $\tilde h_a=h_{d,0}$ and
-$\tilde w_a=\tilde w_{d,0}=\tr\p{\m m_{d,0}^\dag \m w}/N$ for all
-$\abs{d}=a$, and
-\begin{align}
-  \tilde h_{ab} \equiv \f1{\sqrt{\N_a\N_b}}
-  \sum_{\substack{c,d\in\ZZ_L^D\\\abs{c}=a\\\abs{d}=b}} h_{cd}.
-\end{align}
-By rough accounting, isotropy in $D$ dimensions implies one reflection
-symmetry for each of $D$ axes, so the matrix $\ut{\check{\m h}}_0$ has
-dimensions $\sim\p{N/2^D}\times\p{N/2^D}$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The fully collective model}
-
-Here we solve the single- and two-body eigenvalue problems in the case
-of a fully collective spin model with uniform couplings:
-$h_{pq}=-1+\delta_{pq}$, where $\delta_{pq}=1$ if $p=q$ and zero
-otherwise.  In this case, the single-body excitation energies
-$\Delta_k$ are
-\begin{align}
-  \Delta_k
-  = - \sum_{d\ne 0} \sp{\cos\p{d\c k}-1}
-  = N \p{1-\delta_{k,0}}.
-\end{align}
-In order to solve the two-body eigenvalue problem, we expand
-\begin{align}
-  \ut{\check{\m h}}_k
-  &= -\sum_{\substack{c,d\ne0\\c\ne d}}
-  \p{1+e^{-i\p{d-c}\c k}} \op{d}{c}
-  + \sum_{d\ne0} \sp{2\p{N-1}-2} \op{d} \\
-  &= -\sum_{c,d\ne0} \p{1+e^{-i\p{d-c}\c k}} \op{d}{c}
-  + 2\p{N-1} \1_{N-1},
-\end{align}
-where $\1_{N-1}\equiv\sum_{d\ne0}\op{d}$ is the $\p{N-1}\times\p{N-1}$
-identity matrix, and define the vectors
-\begin{align}
-  \ket{\ut{p}}
-  \equiv \f1{\sqrt{N-1}} \sum_{d\ne 0} e^{i p\c d}\ket{d},
-\end{align}
-in terms of which
-\begin{align}
-  \ut{\check{\m h}}_k
-  = 2 \p{N-1} \1_{N-1} - \p{N-1} \p{\op{\ut{0}} + \op{\ut{k}}}.
-\end{align}
-The vectors $\ket{\ut{p}}$ are normalized, but not orthogonal:
-$\braket{\ut{p}|\ut{q}}=\delta_{pq}-\p{1-\delta_{pq}}/\p{N-1}$.  We
-therefore have
-\begin{align}
-  \ut{\check{\m h}}_k
-  = 2 \p{N-1} \1_{N-1}
-  - \sum_{s\in\set{\pm}} \varepsilon_s \op{\ut{k}^s}
-\end{align}
-where
-\begin{align}
-  \varepsilon_\pm \equiv N - 1 \mp 1,
-  &&
-  \ket{\ut{k}^\pm}
-  \equiv \p{1\mp\f1{N-1}}^{-1/2}
-  \f{\ket{\ut{0}}\pm\ket{\ut{k}}}{\sqrt{2}}.
-\end{align}
-The vectors $\ket{\ut{k}^\pm}$ are orthonormal when $k\ne0$.  When
-$k=0$, meanwhile, we have
-\begin{align}
-  \ut{\check{\m h}}_0 = 2\p{N-1} \p{\1_{N-1} - \op{\ut{0}}}.
-\end{align}
-Remembering from \eqref{eq:cycle_vecs} that vectors $\ket{\ut{k}^\pm}$
-are associated with the cycle matrices $\m m_{dk}$ defined in
-\eqref{eq:cycle} via
-\begin{align}
-  \ket{\ut{k}^\pm}
-  \propto \ket{\ut{0}} \pm \ket{\ut{k}}
-  \propto \sum_{d\ne0} \p{1 \pm e^{ik\c d}} \ket{d}
-  \to \sum_{d\ne0} \p{1 \pm e^{ik\c d}} \m m_{dk},
-\end{align}
-one can work out that the vectors $\ket{\ut{k}^\pm}$ correspond to
-matrices that are strictly symmetric ($+$) or anti-symmetric ($-$)
-under transposition.  The two-body perturbations considered in this
-work, however, are defined using coefficient vectors
-$\set{\v{\m w}_O}$ that correspond to symmetric matrices
-$\set{\m w_O}$, with $w_{Opq}=w_{Oqp}$.  We can therefore neglect the
-eigenvectors $\ket{\ut{k}^-}$ and corresponding eigenvalues
-$\varepsilon_-$ entirely for the purposes of this work.
-
-Altogether, the excitation energies corresponding to non-trivial
-eigenstates of $\ut{\check{\m h}}_k$ are
-\begin{align}
-  \Delta_0 \equiv 0,
-  &&
-  \Delta_+ \equiv 2\p{N-1} - \varepsilon_+ = N,
-\end{align}
-while the energy of all orthogonal two-body excitations is
-\begin{align}
-  \Delta_\perp \equiv 2\p{N-1}.
-\end{align}
 
@@
 
-% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \section{Collective and multi-local operators in the permutationally
-%   symmetric manifold}
-% \label{sec:multi_to_collective}
-
-% Here we provide a decomposition multi-local operators into sums of
-% collective operators (and vice versa) within the permutationally
-% symmetric manifold of $N$ spins.  Multi-local operators are operators
-% of the form
-% \begin{align}
-%   O_S \equiv \bigotimes_{O\in S} O,
-%   \label{eq:multi_op_set}
-% \end{align}
-% where $S\equiv\set{O}$ is a set of single-spin operators $O$.  The
-% order of tensor factors in \eqref{eq:multi_op_set} does not matter
-% after projecting operators onto the permutationally symmetric
-% manifold, as we always will in this section.  While the decomposition
-% of a collective operator into a sum of multi-local operators is
-% unique, the opposite is not true: there is generally no unique
-% decomposition of a multi-local operator into a sum of collective
-% operators.  We can, however, enforce a canonical decomposition by
-% choosing a set of collective operators that obey the same symmetries
-% as $O_S$, namely invariance under arbitrary permutations of the
-% single-spin operators $O\in S$.
-
-% Even though $S$ contains $\abs{S}$ spins, a collective operator built
-% from the single-spin operators $O\in S$ may have multiple
-% (single-spin) operators acting on the same spin, resulting in a
-% $k$-local operator with $k<\abs{S}$.  We therefore {\it
-%   partition}\footnote{A partition $P\equiv\set{p}$ of $S$ is a set of
-%   mutually disjoint, non-empty subsets $p\subset S$ with
-%   $\bigcup_{p\in P}p=S$.} $S$ into subsets of operators that act on
-% the same spin.  For any partition $P\equiv\set{p}$ of $S$ (here
-% $p\subset S$ is a subset of operators $O\in S$ that address the same
-% spin), we then define the $\abs{P}$-local operator
-% \begin{align}
-%   O_P \equiv \bigotimes_{p\in P} \bigodot_{O\in p} O,
-% \end{align}
-% where $\bigodot_{x\in\X} x$ denotes a symmetrized product of all
-% elements in $\X$, i.e.
-% \begin{align}
-%   \bigodot_{x\in\X} x
-%   \equiv \f1{\abs{\X}!}
-%   \sum_{\substack{\t{permutations}\\\sigma~\t{of}~\X}}
-%   ~ \prod_{k=1}^{\abs{\X}} x_{\sigma_k},
-% \end{align}
-% with $\sigma_k$ the $k$-th element of the permutation $\sigma$ when
-% expressed in one-line notation; for example,
-% \begin{align}
-%   \bigodot_{j=1}^2 x_j &= x_1 \odot x_2 = \f12 \p{x_1x_2 + x_2x_1}, \\
-%   \bigodot_{j=1}^3 x_j
-%   &= \f1{3!} \p{x_1x_2x_3 + x_1x_3x_2 + x_2x_1x_3
-%     + x_2x_3x_1 + x_3x_1x_2 + x_3x_2x_1}.
-% \end{align}
-% The set of multi-local operators $\set{O_P:\t{partitions}~P~\t{of}~S}$
-% consists of all operators that
-% \begin{enumerate*}
-% \item can be built from ordinary products and tensor products of all
-%   single-spin operators $O\in S$, and
-% \item are invariant under arbitrary permutations of $S$.
-% \end{enumerate*}
-% For any set $S$ of single-spin operators and any partition $P$ of $S$,
-% we can similarly define the collective operators
-% \begin{align}
-%   \O_S \equiv \bigodot_{O\in S} \col{O},
-%   &&
-%   \O_P \equiv \bigodot_{p\in P} \col{\textstyle\bigodot_{O\in p} O},
-% \end{align}
-% where we underline a single-spin operator to denote its collective
-% version, i.e.~$\col{Q}\equiv\sum_j Q^{(j)}$, where $Q^{(j)}$ is an
-% operator that acts with the single-spin operator $Q$ on spin $j$ and
-% trivially on all other spins.  By construction, the set of collective
-% operators $\set{\O_P:\t{partitions}~P~\t{of}~S}$ is one-to-one to the
-% set of multi-local operators $\set{O_P:\t{partitions}~P~\t{of}~S}$,
-% and it is straightforward to work out that for any set $S$ of
-% single-spin operators,
-% \begin{align}
-%   \O_S
-%   \EQPS \sum_{\substack{\t{partitions}\\P~\t{of}~S}}
-%   N!_{\abs{P}} O_P,
-%   &&
-%   N!_k \equiv \prod_{j=0}^{k-1}\p{N-j}! = \f{N!}{\p{N-k}!},
-%   \label{eq:collective_to_multi_full}
-% \end{align}
-% where $\EQPS$ denotes equality under a restriction to the fully
-% symmetric manifold.  For any multi-local operator $O_S$ indexed by a
-% set $S$ of single-spin operators, we can collect expansions of the
-% form \eqref{eq:collective_to_multi_full} for all collective operators
-% built from non-empty subsets of $S$ into a system of linear equations
-% $\L$, and in turn solve $\L$ for $O_S$.  When
-% $\abs{S}\in\set{1,2,3,4}$ (and, by conjecture, for all $S$), by use of
-% a computer algebra system (namely, Mathematica) we find that
-% \begin{align}
-%   N!_{\abs{S}} O_S
-%   &\EQPS \sum_{\substack{\t{partitions}\\P~\t{of}~S}}
-%   \p{-1}^{\abs{S}-\abs{P}} \O_P \times \prod_{p\in P}\p{\abs{p}-1}!
-%   \label{eq:multi_to_collective_full} \\
-%   &\EQPS \f1{\abs{S}!} \sum_{\substack{\t{permutations}\\\Pi~\t{of}~S}}
-%   \sum_{\substack{\t{ordered}\\\t{partitions}\\\tilde P~\t{of}~\Pi}}
-%   \p{-1}^{\abs{S}-\abs{\tilde P}} \O_{\tilde P}
-%   \times \prod_{\tilde p\in\tilde P} \p{\abs{\tilde p}-1}!
-%   \label{eq:multi_to_collective_full_simp}
-% \end{align}
-% where we identify permutations $\Pi$ of $S$ with an ordered list
-% (tuple) $\p{O_1,O_2,\cdots,O_{\abs{S}}}$ of all single-spin operators
-% in $S$; and we enforce that partitions $\tilde P$ of an ordered list
-% $\Pi$, as well as parts $\tilde p\in\tilde P$, are lexicographically
-% ordered in accordance with the ordering of operators in $\Pi$.  If
-% $S=\set{X,Y,Z}$, for example, then the ordered partitions of the
-% trivial permutation $\Pi=\p{X,Y,Z}$ are
-% \begin{align}
-%   \p{\p{X,Y,Z}}, &&
-%   \p{\p{X,Y},\p{Z}}, &&
-%   \p{\p{X,Z},\p{Y}}, &&
-%   \p{\p{X},\p{Y,Z}}, &&
-%   \p{\p{X},\p{Y},\p{Z}}.
-% \end{align}
-% For ordered lists $\tilde S$ and ordered partitions $\tilde P$, we
-% define
-% \begin{align}
-%   O_{\tilde S} \equiv \bigotimes_{O\in\tilde S} O,
-%   &&
-%   O_{\tilde P}
-%   \equiv \bigotimes_{\tilde p\in\tilde P} \prod_{O\in\tilde p} O,
-%   &&
-%   \O_{\tilde S} \equiv \prod_{O\in\tilde S} \col{O},
-%   &&
-%   \O_{\tilde P}
-%   \equiv \prod_{\tilde p\in\tilde P}
-%   \col{\textstyle\prod_{O\in\tilde p} O},
-% \end{align}
-% where products over elements of $\tilde S,\tilde P,\tilde p$ are taken
-% in order according to the elements of these ordered lists, e.g.~if
-% $\tilde S=\p{O_1,O_2,\cdots,O_{\abs{S}}}$, then
-% $\prod_{O\in\tilde S}\equiv O_1O_2\cdots O_{\abs{S}}$.  These
-% definitions allow us to expand
-% \begin{align}
-%   \O_{\tilde S} \EQPS
-%   \sum_{\substack{\t{ordered}\\\t{partitions}\\\tilde P~\t{of}~\tilde S}}
-%   N!_{\abs{\tilde P}} O_{\tilde P},
-% \end{align}
-% in analogy to \eqref{eq:collective_to_multi_full}.
-
-% While the expansions in \eqref{eq:multi_to_collective_full} and
-% \eqref{eq:multi_to_collective_full_simp} closely resemble each other,
-% in practice the latter involves less computational overhead due to the
-% lack of symmetrization in the construction of $\O_{\tilde P}$.  There
-% are, in principle, even more economical decompositions of a
-% multi-local operators into collective opreators, for example
-% \begin{align}
-%   N!_2 X\otimes Y
-%   &\EQPS \col{X}\,\col{Y} - \col{XY}, \\
-%   N!_3 X\otimes Y \otimes Z
-%   &\EQPS \col{X}\,\col{Y}\,\col{Z}
-%   - \col{X}\,\col{YZ} - \col{Y}\,\col{XZ} - \col{Z}\,\col{XY}
-%   + \col{ZXY} + \col{YXZ},
-% \end{align}
-% but these decompositions generally do not preserve the manifest
-% symmetry of a multi-local operator $O_S$ under arbitrary permutations
-% of $S$, and we do not know of any general prescription to arrive at
-% such decompositions aside from solving systems of equations relating
-% collective operators to multi-local operators.
-
-% On a final note, we consider the special case that all operators
-% $O\in S$ are identical, for which we provide the expansions
-% \begin{align}
-%   N!_2 X\otimes X \EQPS \col{X}^2 - \col{X^2},
-%   &&
-%   N!_3 X^{\otimes 3} \EQPS \col{X}^3
-%   - 3 \col{X} \odot \col{X^2} + 2\col{X^3},
-% \end{align}
-% \begin{align}
-%   N!_4 X^{\otimes 4}
-%   \EQPS \col{X}^4 - 4 \col{X}^2 \odot \col{X^2}
-%   - 2 \col{X} \, \col{X}^2 \, \col{X}
-%   + 3 \p{\col{X^2}}^2 + 8 \col{X} \odot \col{X^3}
-%   - 6 \col{X^4}.
-% \end{align}
-% If furthermore $X^2=1$, as with e.g.~the Pauli operators of SU(2),
-% then
-% \begin{align}
-%   N!_2 X\otimes X \EQPS \col{X}^2 - N,
-%   &&
-%   N!_3 X^{\otimes 3} \EQPS \col{X}^3 - \p{3N-2} \col{X},
-% \end{align}
-% \begin{align}
-%   N!_4 X^{\otimes 4} \EQPS \col{X}^4
-%   - 2\p{3N-4} \col{X}^2 + 3 N \p{N-2}.
-% \end{align}
-
 \bibliography{multilevel_spin_notes.bib}

### 2020-03-05 collective_sun/sun_perturbation.tex
a bit of cleanup of the multi-body eigenvalue problem with TI

@@
 $\m x*\m y$ denotes an element-wise (Kronecker) product of $\m x$ and
-$\m y$, with entries $x_ky_k$.
-
-The remaining terms in \eqref{eq:H_VM_psi_start} with only one of
-$p\in k$ or $q\in k$ are
+$\m y$, with entries $x_ky_k$.  The remaining terms in
+\eqref{eq:H_VM_psi_start} with only one of $p\in k$ or $q\in k$ are
 \begin{align}
@@
   = \sum_{k\in\C_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}}
-  \sum_{j=1}^M h_{p k_j} w_k \Pi_{p k_j} O_k \ket\psi.
-\end{align}
-Defining
-$k_{jp}\equiv\p{k_1,k_2,\cdots,k_{j-1},p,k_{j+1},\cdots,k_M}$, i.e.~an
-element of $\C_N\p{M}$ in which the $j$-th number $k_j$ has been
+  \sum_{a=1}^M h_{p k_a} w_k \Pi_{p k_a} O_k \ket\psi.
+\end{align}
+where $k\equiv\p{k_1,k_2,\cdots,k_M}$, with $k_a$ the $a$-th entry in
+$k$.  Defining
+$k_{ap}\equiv\p{k_1,k_2,\cdots,k_{a-1},p,k_{a+1},\cdots,k_M}$, i.e.~an
+element of $\C_N\p{M}$ in which the $a$-th number $k_a$ has been
 replaced by $p$, we can simplify
 \begin{align}
-  \Pi_{p k_j} O_k \ket\psi
-  = \Pi_{p k_j} O_k \Pi_{p k_j} \ket\psi
-  = O_{k_{jp}}\ket\psi,
-\end{align}
-and switch the labels for $p$ and $k_j$ to find that
+  \Pi_{p k_a} O_k \ket\psi
+  = \Pi_{p k_a} O_k \Pi_{p k_a} \ket\psi
+  = O_{k_{ap}}\ket\psi,
+\end{align}
+and switch the labels for $p$ and $k_a$ to find that
 \begin{align}
@@
   = \sum_{k\in\C_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}}
-  \sum_{j=1}^M h_{pk_j} w_{k_{jp}} O_k \ket\psi.
-\end{align}
-For each fixed $j$, the sum over $p$ above is essentially a
-contraction of the matrix $\m h$ with the $j$-th index of the tensor
-$\m w$.  Denoting such a contraction by $\m h\c_j\m w$, we can
+  \sum_{a=1}^M h_{pk_a} w_{k_{ap}} O_k \ket\psi.
+\end{align}
+For each fixed $a$, the sum over $p$ above is essentially a
+contraction of the matrix $\m h$ with the $a$-th index of the tensor
+$\m w$.  Denoting such a contraction by $\m h\c_a\m w$, we can
 therefore write this result as
@@
   &&
-  \m h\circ\m w \equiv \sum_{j=1}^M \m h\c_j\m w,
+  \m h\circ\m w \equiv \sum_{a=1}^M \m h\c_a\m w,
 \end{align}
@@
   \p{\m h\circ\m w}_k
-  \equiv \sum_{\substack{p\in\ZZ_N\\p\notin k}} \sum_{j=1}^M
-  h_{pk_j} w_{k_{jp}}
+  \equiv \sum_{\substack{p\in\ZZ_N\\p\notin k}} \sum_{a=1}^M
+  h_{pk_a} w_{k_{ap}}
   = \sum_{\substack{p\in\ZZ_N\\p\notin k}}
@@
 symmetry and translational invariance.  To this end, we first expand
-the coupling tensor $\m w$ in a complete basis for all tensors of
-dimension $M$:
-\begin{align}
-  \m w = \sum_{k\in\ZZ_N^M} w_k \m e_k,
-\end{align}
-where $\m e_k$ is a unit tensor with one in the entry indexed by $k$
-and zeros everywhere else.  We then use the permutational symmetry of
-$\m w$ to expand it in a basis for symmetric tensors, making use of
-the fact that all entries of $\m w$ with repeated indices are zero:
+the coupling tensor $\m w$ in a complete basis for symmetric tensors:
 \begin{align}
@@
 \end{align}
-where $P_M$ is the set of all permutations of $M$ symbols.  On a
-$D$-dimensional lattice of linear size $L$, translational invariance
-implies that the coefficients $w_k$ remain the same if all spins
-indexed by $k$ are shifted by the same displacement $d\in\ZZ_L^D$.  To
-account for this invariance, we denote a translation of all spins in
-$k$ by a displacement $d\in\ZZ_L^D$ by $k_d$, and define the
-translation-symmetrized basis tensors
+where $P_M$ is the set of all permutations of $M$ symbols, and
+$\m e_\ell$ is a unit tensor with one in the entry indexed by $\ell$
+and zeros everywhere else.  On a $D$-dimensional lattice of linear
+size $L$, translational invariance implies that the coefficients $w_k$
+remain the same if all spins indexed by $k$ are shifted by the same
+displacement $d\in\ZZ_L^D$.  To account for this invariance, we denote
+a translation of all spins in $k$ by a displacement $d\in\ZZ_L^D$ by
+$k_d$, and define the translation-symmetrized basis tensors
 \begin{align}
@@
 otherwise might not occur if some translations of $k$ results in the
-same choice of spins.  We then define the set $\T_N\p{M}$ of
-equivalence classes in $\C_N\p{M}$ under arbitrary translations, and
-expand
-\begin{align}
-  \m w = \sum_{k\in\T_N\p{M}} w_k \m m_k^{\t{TS}}.
+same choice of spins.  We then define the set $\T_{LD}\p{M}$ of
+equivalence classes in $\C_N\p{M}$ under arbitrary translations on
+$D$-dimensional lattice of linear size $L$, and expand
+\begin{align}
+  \m w = \sum_{k\in\T_{LD}\p{M}} w_k \m m_k^{\t{TS}}.
   \label{eq:sym_shift_tensor}
@@
   \m h\circ\m w
-  = \sum_{k\in\T_N\p{M}} \m m_k^{\t{TS}}
-  \sum_{d\in\ZZ_L^D} h_{d,0} \sum_{a=1}^M w_{k_{d;a}},
+  = \sum_{k\in\T_{LD}\p{M}} \sum_{d\in\ZZ_L^D} \sum_{a=1}^M w_{k_{d;a}}
+  h_{d,0} w_{k_{d;a}} \m m_k^{\t{TS}},
 \end{align}
@@
 \begin{align}
+  \bar{\m h}
+  = \sum_{k\in\T_{LD}\p{M}} \sp{M h_0 - \sum_{p,q\in k} h_{pq}}
+  \m m_k^{\t{TS}},
+  &&
+  h_0 \equiv \sum_{d\in\ZZ_L^D} h_{d,0},
+\end{align}
+which implies that
+\begin{align}
   \bar{\m h} * \m w = M h_0 \m w
-  - \sum_{k\in\T_N\p{M}} w_k \m m_k^{\t{TS}}
-  \sum_{p,q\in k} h_{pq}
-  &&
-  h_0 \equiv \sum_{d\in\ZZ_L^D} h_{d,0}.
+  - \sum_{k\in\T_{LD}\p{M}} \sum_{p,q\in k} h_{pq} w_k \m m_k^{\t{TS}}.
 \end{align}
@@
 \begin{align}
-  \sum_{k\in\T_N\p{M}} \m m_k^{\t{TS}}
-  \sp{\sum_{p,q\in k} h_{pq} w_k
+  \sum_{k\in\T_{LD}\p{M}} \sp{\sum_{p,q\in k} h_{pq} w_k
     + \sum_{d\in\ZZ_L^D} h_{d,0} \sum_{a=1}^M w_{k_{d;a}}}
+  \m m_k^{\t{TS}}
   = \p{M h_0 + \Delta} \m w,
@@
   \ut{\check{\m h}} \equiv
-  \sum_{k\in\T_N\p{M}} \ket{k}
+  \sum_{k\in\T_{LD}\p{M}} \ket{k}
   \sp{\sum_{p,q\in k} h_{pq} \bra{k}
@@
   &&
-  \v{\m w} \equiv \sum_{k\in\T_N\p{M}} w_k \ket{k},
+  \v{\m w} \equiv \sum_{k\in\T_{LD}\p{M}} w_k \ket{k},
 \end{align}
@@
 $k$ are the same up to a global translation.
-
-\vspace{1cm}
-
-\red{TODO: write codes to check above work, fix any errors, explicitly
-  write out two-body reduced eigenvalue problems, include sub-section
-  about isotropy, replace the following appendix}
-
-\vspace{1cm}
 

### 2020-03-05 collective_sun/sun_perturbation.tex
added section on computing expectation values of multi-local operators

@@
 
+\newcommand{\A}{\mathcal{A}}
 \newcommand{\C}{\mathcal{C}}
@@
 \begin{align}
-  {M_j \choose a_1,a_2,\cdots,a_m}
-  \equiv \f{M_j!}{a_1!a_2!\cdots a_m!}
+  \C\p{a_1,a_2,\cdots,a_m}
+  \equiv \f{\p{a_1+a_2+\cdots+a_m}!}{a_1!a_2!\cdots a_m!}
 \end{align}
@@
 
-Here we discuss the rules governing symmetry factors that appear when
-decomposing diagrams.  In general, eliminating a filled dot from a
-region with $F$ filled dots makes a diagram pick up a factor of $1/F$,
-e.g.
+As seen in \eqref{eq:example_elim_final}, the process of simplifying
+diagrams generally involves symmetry factors that need to be kept
+track of by hand.  Here we discuss the rules governing symmetry
+factors that appear when decomposing diagrams.  In general,
+eliminating a filled dot from a region with $F$ filled dots makes a
+diagram pick up a factor of $1/F$, e.g.
 \begin{align}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Matrix elements of multi-body operators}
-
-\red{TODO: write section}
+\subsubsection{Matrix elements of individual multi-spin operators}
+
+The process of constructing and simplifying diagrams to compute the
+coefficients $w\p{g}$ of the diagrammatic operator product expansion
+in \eqref{eq:sym_prod_group} is systematic enough to be converted into
+an algorithm for execution on a computer.  The last step necessary to
+automate the evaluation of \eqref{eq:sym_prod_group} is expansion of
+the multi-spin operators $O\p{g}$ in a basis for the permutationally
+symmetric manifold $\M_0$.  Without loss of generality, we consider a
+single permutationally symmetric $M$-spin operator $Q$; our task is
+essentially to find coefficients of the expansion
+\begin{align}
+  \P_0 Q \P_0
+  = \sum_{a,b} \bk{a|Q|b} \op{a}{b},
+\end{align}
+where $\set{\ket{a}},\set{\ket{b}}$ are complete bases for $\M_0$, and
+the choice of $M$ spins on which $Q$ acts is arbitrary.  In a system
+of $N$ spins with $n$ states each, a suitable basis for $\M_0$ can be
+labeled by the occupation number of each single-spin state, namely
+$a=\p{a_1,a_2,\cdots,a_n}$ with each $a_j\ge 0$ and $\sum_ja_j=N$.
+Written out explicitly,
+\begin{align}
+  \ket{a} = \f1{\sqrt{\C\p{a}}}
+  \sum_{\substack{\t{distinct}\\\t{permutations}\\\Pi~\t{of}~\tilde a}}
+  \Pi \ket{\tilde a},
+  &&
+  \C\p{a} \equiv \f{\p{\sum_j a_j}!}{\prod_ja_j!}
+  &&
+  \ket{\tilde a} \equiv \bigotimes_{j=1}^n \ket{j}^{\otimes a_j},
+\end{align}
+where the multinomial coefficient $\C\p{a}$ counts the number of
+distinct ways to permute the tensor factors of $\ket{\tilde a}$, and
+enforces $\bk{a|a}=1$.  We denote the set of all valid assignments of
+$S$ spins to $n$ states by $\A_n\p{S}$, such that the standard basis
+for the permutationally symmetric manifold of $S$ spins is
+$\set{\ket{a}:a\in\A_n\p{S}}$.  Identifying this standard basis allows
+us to expand
+\begin{align}
+  \bk{a|Q|b} = \sum_{\alpha,\beta\in\A_n\p{M}}
+  \delta_{a-\alpha,b-\beta}
+  \sqrt{\C\p{a-\alpha}\C\p{b-\beta}}
+  \sqrt{\f{\C\p{\alpha}\C\p{\beta}}{\C\p{a}\C\p{b}}}
+  \bk{\alpha|Q|\beta},
+  \label{eq:multi_body_eval}
+\end{align}
+where the differences $a-\alpha$ and $b-\beta$ are understood to be
+taken element-wise; $\delta_{a-\alpha,b-\beta}=1$ if
+$a-\alpha=b-\beta$ and zero otherwise; and for ease of notation we
+define $\C\p{c}\equiv 0$ if any entries of $c$ are negative.  We sum
+over both $\alpha$ and $\beta$ above merely to keep the expression
+symmetric with respect to transposition; in practice, one can simply
+sum over $\alpha\in\A_n\p{M}$ and set $\beta\equiv b-a+\alpha$.  Note
+that, by slight abuse of notation, the operator $Q$ on the left of
+\eqref{eq:multi_body_eval} acts on an arbitrary choice of $M$ spins
+(out of $N$), whereas the operator $Q$ on the right of
+\eqref{eq:multi_body_eval} is simply an $M$-spin operator, with matrix
+elements $\bk{\alpha|Q|\beta}$ evaluated with respect to the $M$-spin
+states $\ket\alpha,\ket\beta$.
 

### 2020-03-05 collective_sun/sun_perturbation.tex
clarified an association between an operator and its restriction to a subspace
minor change in notation

@@
 \begin{align}
-  \check{\m h} \c \v{\m w} = \Delta \v{\m w},
+  \check{\m h}_2 \c \v{\m w} = \Delta \v{\m w},
   \label{eq:V2_eig}
@@
 \begin{align}
-  \check{\m h}
+  \check{\m h}_2
   \equiv \sum_{p<q} \ket{\set{p,q}} \sp{\sum_{k\notin\set{p,q}}
@@
 ${N \choose 2}\times{N \choose 2}\sim N^2\times N^2$ matrix
-$\check{\m h}$.  When both he interaction Hamiltonian $H_0$
+$\check{\m h}_2$.  When both he interaction Hamiltonian $H_0$
 translationally invariant, however, this eigenvalue problem can be
@@
 of $\v{\m w}$ onto the $\Delta$-eigenspace of the matrix
-$\check{\m h}$ defined in \eqref{eq:V2_eig_mat} by $\v{\m w}_\Delta$,
-and define
+$\check{\m h}_2$ defined in \eqref{eq:V2_eig_mat} by
+$\v{\m w}_\Delta$, and define
 \begin{align}
@@
 \begin{align}
-  \ut{\check{\m h}} \c \v{\m w} = \p{M h_0 + \Delta} \v{\m w},
+  \check{\m h} \c \v{\m w} = \p{M h_0 + \Delta} \v{\m w},
   \label{eq:H_VM_eig_simp}
@@
 \begin{align}
-  \ut{\check{\m h}} \equiv
+  \check{\m h} \equiv
   \sum_{k\in\T_{LD}\p{M}} \ket{k}
@@
 \begin{align}
-  \Q \simeq_M \P_M \Q \P_M,
+  \Q =_M \P_M \Q \P_M,
   &&
@@
 \end{align}
-where finding matrix elements of $\Q$ now requires computing products
-of the form
+where $=_M$ denotes equality up to a restriction to the support of
+$\P_M$, and finding matrix elements of $\Q$ now requires computing
+products of the form
 \begin{align}

### 2020-03-05 collective_sun/sun_perturbation.tex
clarified size of matrix to diagonalize for the multi-body eigenvalue problem

@@
 $M$-body perturbations $\V_M$ that generate states of definite
-interaction energy nominally requires diagonalizing an
-${N\choose M}\times{N\choose M}\sim N^M\times N^M$ matrix.  Here, we
-simplify this problem to that of diagonalizing an
+interaction energy nominally requires diagonalizing a matrix with
+dimensions ${N\choose M}\times{N\choose M}\sim N^M\times N^M$.  Here,
+we simplify this problem to that of diagonalizing a matrix with
+dimensions
 $\sim{N-1\choose M-1}\times{N-1\choose M-1}\sim N^{M-1}\times N^{M-1}$
-matrix when both the interaction Hamiltonian $H_0$ and the
-perturbation $\V_M$ are translationally invariant.  This
-simplification reduces the computational cost of solving the $M$-body
-eigenvalue problem by a factor of $\O\p{N^3}$.
+when both the interaction Hamiltonian $H_0$ and the perturbation
+$\V_M$ are translationally invariant.  This simplification reduces the
+computational cost of solving the $M$-body eigenvalue problem by a
+factor of $\O\p{N^3}$.
 
@@
 repeated values, and enforce $\ket{j}=\ket{k}$ if spins $j$ and spins
-$k$ are the same up to a global translation.
+$k$ are the same up to a global translation.  The matrix
+$\check{\m h}$ has dimensions
+$\abs{\T_{LD}\p{M}}\times\abs{\T_{LD}\p{M}}$, where
+$\abs{\T_{LD}\p{M}} \approx { N-1 \choose M-1 }/M \sim N^{M-1}$.
 

### 2020-03-06 collective_sun/sun_perturbation.tex
clarified that the spectral gap occurs with long-range interactions
added section on simplifying the multi-body eigenvalue problem further in the presence of additional symmetries (such as isotropy)

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Existence of a spectral gap with power-law interactions}
+\section{Existence of a spectral gap with long-range power-law
+  interactions}
 \label{sec:spectral_gap}
@@
 $\abs{\T_{LD}\p{M}} \approx { N-1 \choose M-1 }/M \sim N^{M-1}$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Additional symmetries}
+
+If the interaction Hamiltonian $H_0$ and multi-body perturbation
+$\V_M$ have additional symmetries, such as isotropy or nonchirality,
+then we can use this symmetry to reduce the eigenvalue problem in
+\eqref{eq:H_VM_eig_simp} even further.  To this end, we define the set
+$\S_{LD}\p{M}$ of equivalence classes in $\T_{LD}\p{M}$ under the
+relevant symmetry.  In the case of isotropic two-body perturbations,
+for example, the set of translationally-symmetrized spin choices
+$\T_{LD}\p{2}$ can be classified by {\it displacements} on a lattice
+(i.e.~a vector from one chosen spin to the other), and we can equate
+all displacements that have the same magnitude, so elements of
+$\S_{LD}\p{2}$ can be classified by all by {\it distances} between any
+two points on the lattice.  For any member $s\in\S_{LD}\p{M}$, we then
+define a uniform superposition $\ket{s}$ over all corresponding
+members of $\T_{LD}\p{M}$, as well as the projector $\P_\S$ onto the
+space of uniform superpositions:
+\begin{align}
+  \ket{s} \equiv \f1{\sqrt{\abs{s}}} \sum_{k\in s} \ket{k},
+  &&
+  \P_\S \equiv \sum_{s\in\S_{LD}\p{M}} \op{s},
+\end{align}
+where $\abs{s}$ is the number of members of $\T_{LD}\p{M}$ in the
+equivalence class $s$.  We then expand the (symmetric) coefficient
+vector
+\begin{align}
+  \v{\m w} = \P_\S \v{\m w}
+  = \sum_{s\in\S_{LD}\p{M}} \ut{w}_s \ket{s},
+  &&
+  \ut{w}_s \equiv w_s \sqrt{\abs{s}},
+\end{align}
+as well as the symmetric projection of the matrix $\check{\m h}$:
+\begin{align}
+  \P_\S \check{\m h} \P_\S
+  = \sum_{s\in\S_{LD}\p{M}} \ket{s} \sp{\f{h_s}{\abs{s}} \bra{s}
+    + \sum_{d\in\ZZ_L^D} \f{h_{d,0}}{\sqrt{\abs{s}}} \sum_{a=1}^M
+    \f1{\sqrt{\abs{s_{d;a}}}} \bra{s_{d;a}}},
+\end{align}
+where $h_s\equiv\sum_{p,q\in k}h_{pq}$ for any $k\in s$; and,
+enforcing a canonical choice of $k\in s$ for all $s\in\S_{LD}\p{M}$,
+$s_{d;a}$ is the equivalence class of the corresponding $k_{d;a}$,
+with $\ket{s_{d;a}}\equiv0$ if $\ket{k_{d;a}}=0$.  The
+symmetry-reduced eigenvalue problem is then
+\begin{align}
+  \P_\S \check{\m h} \P_\S \c \v{\m w} = \p{M h_0 + \Delta} \v{\m w}.
+\end{align}
 
@@
   \delta_{a-\alpha,b-\beta}
-  \sqrt{\C\p{a-\alpha}\C\p{b-\beta}}
-  \sqrt{\f{\C\p{\alpha}\C\p{\beta}}{\C\p{a}\C\p{b}}}
+  \sqrt{\f{\C\p{\alpha}\C\p{a-\alpha}\C\p{\beta}\C\p{b-\beta}}
+    {\C\p{a}\C\p{b}}}
   \bk{\alpha|Q|\beta},

### 2020-03-06 collective_sun/sun_perturbation.tex
minor change in language/notation in the section on evaluating multi-spin operators

@@
 \begin{align}
-  \bk{a|Q|b} = \sum_{\alpha,\beta\in\A_n\p{M}}
+  \bk{a|Q|b}
+  = \sum_{\substack{\alpha,\beta\in\A_n\p{M}\\\alpha\le a,\,\beta\le b}}
   \delta_{a-\alpha,b-\beta}
@@
 \end{align}
-where the differences $a-\alpha$ and $b-\beta$ are understood to be
-taken element-wise; $\delta_{a-\alpha,b-\beta}=1$ if
-$a-\alpha=b-\beta$ and zero otherwise; and for ease of notation we
-define $\C\p{c}\equiv 0$ if any entries of $c$ are negative.  We sum
-over both $\alpha$ and $\beta$ above merely to keep the expression
-symmetric with respect to transposition; in practice, one can simply
-sum over $\alpha\in\A_n\p{M}$ and set $\beta\equiv b-a+\alpha$.  Note
-that, by slight abuse of notation, the operator $Q$ on the left of
-\eqref{eq:multi_body_eval} acts on an arbitrary choice of $M$ spins
-(out of $N$), whereas the operator $Q$ on the right of
-\eqref{eq:multi_body_eval} is simply an $M$-spin operator, with matrix
-elements $\bk{\alpha|Q|\beta}$ evaluated with respect to the $M$-spin
-states $\ket\alpha,\ket\beta$.
+where the restriction $\alpha\le a$ and difference $a-\alpha$ are
+evaluated element-wise, i.e.~$\alpha<a\implies \alpha_j\le a_j$ and
+$\p{a-\alpha}_j=a_j-\alpha_j$ for all $j$; and $\delta_{cd}=1$ if
+$c=d$ and zero otherwise.  We sum over both $\alpha$ and $\beta$ above
+merely to keep the expression symmetric with respect to transposition;
+in practice, one can simply sum over $\alpha\in\A_n\p{M}$ and set
+$\beta\equiv b-a+\alpha$.  Note that, by slight abuse of notation, the
+operator $Q$ on the left of \eqref{eq:multi_body_eval} acts on an
+arbitrary choice of $M$ spins (out of $N$), whereas the operator $Q$
+on the right of \eqref{eq:multi_body_eval} is simply an $M$-spin
+operator, with matrix elements $\bk{\alpha|Q|\beta}$ evaluated with
+respect to the $M$-spin states $\ket\alpha,\ket\beta$.
 

### 2020-03-06 collective_sun/sun_perturbation.tex
refer to appendices as sections

@@
 example, always yield a non-vanishing spectral gap when $\alpha\le D$
-(see Appendix \ref{sec:spectral_gap}).
+(see Section \ref{sec:spectral_gap}).
 
@@
 $H_1^{(2)}$, we first choose an arbitrary permutationally symmetric
-state $\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_V1_psi}):
+state $\ket\psi\in\M_0$ and expand (see Section \ref{sec:H_V1_psi}):
 \begin{align}
@@
 eigenvalue problem can be solved analytically; we provide the
-corresponding solution at the end of Appendix \ref{sec:H_V1_psi}.
+corresponding solution at the end of Section \ref{sec:H_V1_psi}.
 
@@
 where the product $\P_0 \p{\V_1^\Delta}^2 \P_0$ is worked out in
-Appendix \ref{sec:PXYP}.  In total, we find that
+Section \ref{sec:PXYP}.  In total, we find that
 \begin{align}
@@
 $H_2^{(2)}$, as before we pick an arbitrary permutationally symmetric
-state $\ket\psi\in\M_0$ and expand (see Appendix
-\ref{sec:eigenstates})
+state $\ket\psi\in\M_0$ and expand (see Section \ref{sec:eigenstates})
 \begin{align}
@@
 reduced to that of diagonalizing an $\sim N\times N$ matrix; we
-perform this reduction in Appendix \ref{sec:eigenstates_TI}.
+perform this reduction in Section \ref{sec:eigenstates_TI}.
 
@@
 where the product $\P_0 \p{\V_2^\Delta}^2 \P_0$ is simplified in
-Appendix \ref{sec:POQP}.  In the special case that
+Section \ref{sec:POQP}.  In the special case that
 \begin{align}
@@
 $N=L^D$ spins, the spectral gap of the interaction Hamiltonian is (see
-Appendix \ref{sec:H_V1_psi})
+Section \ref{sec:H_V1_psi})
 \begin{align}

### 2020-03-09 collective_sun/sun_perturbation.tex
minor change in notation

@@
   \sp{\sum_k \p{h_{pk} w_{Okq} + h_{qk} w_{Opk}}
-  - \bar h_{pq} w_{Opq}}
+    - \tilde h_{pq} w_{Opq}}
   O_{pq} \ket\psi,
@@
 \begin{align}
-  \bar h_{pq} \equiv h_p + h_q - 2 h_{pq},
+  \tilde h_{pq} \equiv h_p + h_q - 2 h_{pq},
   &&
@@
     \p{h_{pk} \bra{\set{k,q}} + h_{qk} \bra{\set{p,k}}}
-    - \bar h_{pq} \bra{\set{p,q}}}.
+    - \tilde h_{pq} \bra{\set{p,q}}}.
   \label{eq:V2_eig_mat}
@@
 translationally invariant, however, this eigenvalue problem can be
-reduced to that of diagonalizing an $\sim N\times N$ matrix; we
-perform this reduction in Section \ref{sec:eigenstates_TI}.
+reduced to that of diagonalizing a matrix of dimensions
+$\sim N\times N$; we perform this reduction in Section
+\ref{sec:exploiting_symmetries}.
 
@@
   \sum_{\substack{\p{p,q}\in\C_N\p{2}\\p,q\in k~\t{or}~p,q\notin k}} h_{pq}
-  = E_0 - \bar h_k,
-  &&
-  \bar h_k \equiv \sum_{p\in k}\p{h_p - \sum_{q\in k}h_{pq}},
+  = E_0 - \tilde h_k,
+  &&
+  \tilde h_k \equiv \sum_{p\in k}\p{h_p - \sum_{q\in k}h_{pq}},
   &&
@@
 \begin{align}
-  \sum_{k\in\C_N\p{M}} \p{E_0 - \bar h_k} w_k O_k \ket\psi
+  \sum_{k\in\C_N\p{M}} \p{E_0 - \tilde h_k} w_k O_k \ket\psi
   = E_0 \V_M \ket\psi
-  - \sum_{k\in\C_N\p{M}} \p{\bar{\m h} * \m w}_k O_k \ket\psi,
-\end{align}
-where $\bar{\m h}$ is a tensor with entries $\bar h_k$; and
+  - \sum_{k\in\C_N\p{M}} \p{\tilde{\m h} * \m w}_k O_k \ket\psi,
+\end{align}
+where $\tilde{\m h}$ is a tensor with entries $\tilde h_k$; and
 $\m x*\m y$ denotes an element-wise (Kronecker) product of $\m x$ and
@@
   = E_0 \V_M \ket\psi + \sum_{k\in\C_N\p{M}}
-  \p{\m h\circ\m w - \bar{\m h} * \m w}_k O_k \ket\psi.
+  \p{\m h\circ\m w - \tilde{\m h} * \m w}_k O_k \ket\psi.
   \label{eq:H_VM_psi_end}
@@
 \begin{align}
-  \m h\circ\m w - \bar{\m h} * \m w = \Delta \m w.
+  \m h\circ\m w - \tilde{\m h} * \m w = \Delta \m w.
   \label{eq:H_VM_psi_eig}
@@
 If $M=1$, then $\p{\m h\circ\m w}_k=\sum_\ell h_{k\ell}w_\ell$ and
-$\bar h_k=h_k$, so
+$\tilde h_k=h_k$, so
 \begin{align}
@@
   \sp{\sum_{k\in\ZZ_N}\p{h_{pk} w_{kq} + h_{qk} w_{pk}}
-    - \bar h_{pq} w_{pq}}
+    - \tilde h_{pq} w_{pq}}
   O_k \ket\psi,
 \end{align}
-where $\bar h_{pq}=h_p+h_q-2h_{pq}$ with $h_p\equiv\sum_qh_{pq}$.
+where $\tilde h_{pq}=h_p+h_q-2h_{pq}$ with $h_p\equiv\sum_qh_{pq}$.
 
@@
 \subsection{Translational invariance}
-\label{sec:eigenstates_TI}
 
@@
 \begin{align}
-  \bar{\m h}
+  \tilde{\m h}
   = \sum_{k\in\T_{LD}\p{M}} \sp{M h_0 - \sum_{p,q\in k} h_{pq}}
@@
 \begin{align}
-  \bar{\m h} * \m w = M h_0 \m w
+  \tilde{\m h} * \m w = M h_0 \m w
   - \sum_{k\in\T_{LD}\p{M}} \sum_{p,q\in k} h_{pq} w_k \m m_k^{\t{TS}}.

### 2020-03-09 collective_sun/sun_perturbation.tex
simplify section on using translational invariance to simplify the multibody eigenvalue problem

@@
 and translational invariance, we need to express the eigenvalue
-equation \eqref{eq:H_VM_psi_eig} in the same basis.  In order to do
-so, we label spins $p,q$ by the position of the corresponding lattice
-site in $\ZZ_L^D$, and use translational invariance of the couplings
-$\m h$ to expand
-\begin{align}
-  \m h \equiv \sum_{p,q\in\ZZ_L^D} h_{pq} \op{p}{q}
-  = \sum_{q,d\in\ZZ_L^D} h_{q+d,q} \op{q+d}{q}
-  = \sum_{d\in\ZZ_L^D} h_{d,0} \m s_d,
-  &&
-  \m s_d \equiv \sum_{q\in\ZZ_L^D} \op{q+d}{d}.
-\end{align}
-Starting from this form of $\m h$ and the expansion of $\m w$ in
-\eqref{eq:sym_tensor}, we can work out that
+equation \eqref{eq:H_VM_psi_eig} in the same basis.  We therefore
+expand
 \begin{align}
   \m h\circ\m w
-  = \sum_{k\in\T_{LD}\p{M}} \sum_{d\in\ZZ_L^D} \sum_{a=1}^M w_{k_{d;a}}
-  h_{d,0} w_{k_{d;a}} \m m_k^{\t{TS}},
-\end{align}
-where $k_{d;a}$ denotes the spins acquired by displacing the $a$-th
-spin of $k$ by $d$.  With a bit of additional work, we can also
-simplify
-\begin{align}
-  \tilde{\m h}
-  = \sum_{k\in\T_{LD}\p{M}} \sp{M h_0 - \sum_{p,q\in k} h_{pq}}
-  \m m_k^{\t{TS}},
-  &&
-  h_0 \equiv \sum_{d\in\ZZ_L^D} h_{d,0},
-\end{align}
-which implies that
-\begin{align}
-  \tilde{\m h} * \m w = M h_0 \m w
-  - \sum_{k\in\T_{LD}\p{M}} \sum_{p,q\in k} h_{pq} w_k \m m_k^{\t{TS}}.
-\end{align}
-Putting everything together, the generalized eigenvalue problem in
-\eqref{eq:H_VM_psi_eig} can be written in the form
-\begin{align}
-  \sum_{k\in\T_{LD}\p{M}} \sp{\sum_{p,q\in k} h_{pq} w_k
-    + \sum_{d\in\ZZ_L^D} h_{d,0} \sum_{a=1}^M w_{k_{d;a}}}
-  \m m_k^{\t{TS}}
-  = \p{M h_0 + \Delta} \m w,
-\end{align}
-or equivalently
-\begin{align}
-  \check{\m h} \c \v{\m w} = \p{M h_0 + \Delta} \v{\m w},
-  \label{eq:H_VM_eig_simp}
+  = \sum_{k\in\T_{LD}\p{M}} \sum_{p\in\ZZ_L^D} \sum_{a=1}^M
+  h_{pk_a} w_{k_{ap}} \m m_k^{\t{TS}},
+  &&
+  \tilde{\m h} * \m w = \sum_{k\in\T_{LD}\p{M}}
+  \tilde h_k w_k \m m_k^{\t{TS}},
+\end{align}
+which implies that the generalized eigenvalue problem in
+\eqref{eq:H_VM_psi_eig} can be written as
+\begin{align}
+  \check{\m h}\c\v{\m w} = \Delta\v{\m w},
 \end{align}
@@
 \begin{align}
-  \check{\m h} \equiv
-  \sum_{k\in\T_{LD}\p{M}} \ket{k}
-  \sp{\sum_{p,q\in k} h_{pq} \bra{k}
-    + \sum_{d\in\ZZ_L^D} h_{d,0} \sum_{a=1}^M \bra{k_{d;a}}},
+  \check{\m h}
+  \equiv \sum_{k\in\T_{LD}\p{M}} \sum_{p\in\ZZ_L^D} \sum_{a=1}^M
+  h_{pk_a} \op{k}{k_{ap}}
+  - \sum_{k\in\T_{LD}\p{M}} \tilde h_k \op{k},
   &&
@@
 \end{align}
-and, to keep notation simple, we define $\ket{k}=0$ for all $k$ with
-repeated values, and enforce $\ket{j}=\ket{k}$ if spins $j$ and spins
-$k$ are the same up to a global translation.  The matrix
-$\check{\m h}$ has dimensions
+and, by slight abuse of notation we identify an equivalence classes
+$k\in\T_{LD}\p{M}$ with a canonical choice of spins
+$\tilde k\in k\subset\C_N\p{M}$ (such that $k_{ap}=\tilde k_{ap}$,
+with $\tilde k_{ap}$ equal to $\tilde k$ with the $a$-th spin replaced
+by $p$); and we define $\ket{j}=0$ for all $j\in\C_N\p{M}$ with
+repeated values.  The matrix $\check{\m h}$ has dimensions
 $\abs{\T_{LD}\p{M}}\times\abs{\T_{LD}\p{M}}$, where
 $\abs{\T_{LD}\p{M}} \approx { N-1 \choose M-1 }/M \sim N^{M-1}$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Additional symmetries}
-
-If the interaction Hamiltonian $H_0$ and multi-body perturbation
-$\V_M$ have additional symmetries, such as isotropy or nonchirality,
-then we can use this symmetry to reduce the eigenvalue problem in
-\eqref{eq:H_VM_eig_simp} even further.  To this end, we define the set
-$\S_{LD}\p{M}$ of equivalence classes in $\T_{LD}\p{M}$ under the
-relevant symmetry.  In the case of isotropic two-body perturbations,
-for example, the set of translationally-symmetrized spin choices
-$\T_{LD}\p{2}$ can be classified by {\it displacements} on a lattice
-(i.e.~a vector from one chosen spin to the other), and we can equate
-all displacements that have the same magnitude, so elements of
-$\S_{LD}\p{2}$ can be classified by all by {\it distances} between any
-two points on the lattice.  For any member $s\in\S_{LD}\p{M}$, we then
-define a uniform superposition $\ket{s}$ over all corresponding
-members of $\T_{LD}\p{M}$, as well as the projector $\P_\S$ onto the
-space of uniform superpositions:
-\begin{align}
-  \ket{s} \equiv \f1{\sqrt{\abs{s}}} \sum_{k\in s} \ket{k},
-  &&
-  \P_\S \equiv \sum_{s\in\S_{LD}\p{M}} \op{s},
-\end{align}
-where $\abs{s}$ is the number of members of $\T_{LD}\p{M}$ in the
-equivalence class $s$.  We then expand the (symmetric) coefficient
-vector
-\begin{align}
-  \v{\m w} = \P_\S \v{\m w}
-  = \sum_{s\in\S_{LD}\p{M}} \ut{w}_s \ket{s},
-  &&
-  \ut{w}_s \equiv w_s \sqrt{\abs{s}},
-\end{align}
-as well as the symmetric projection of the matrix $\check{\m h}$:
-\begin{align}
-  \P_\S \check{\m h} \P_\S
-  = \sum_{s\in\S_{LD}\p{M}} \ket{s} \sp{\f{h_s}{\abs{s}} \bra{s}
-    + \sum_{d\in\ZZ_L^D} \f{h_{d,0}}{\sqrt{\abs{s}}} \sum_{a=1}^M
-    \f1{\sqrt{\abs{s_{d;a}}}} \bra{s_{d;a}}},
-\end{align}
-where $h_s\equiv\sum_{p,q\in k}h_{pq}$ for any $k\in s$; and,
-enforcing a canonical choice of $k\in s$ for all $s\in\S_{LD}\p{M}$,
-$s_{d;a}$ is the equivalence class of the corresponding $k_{d;a}$,
-with $\ket{s_{d;a}}\equiv0$ if $\ket{k_{d;a}}=0$.  The
-symmetry-reduced eigenvalue problem is then
-\begin{align}
-  \P_\S \check{\m h} \P_\S \c \v{\m w} = \p{M h_0 + \Delta} \v{\m w}.
-\end{align}
 

### 2020-03-09 collective_sun/sun_perturbation.tex
clarified some definitions

@@
   &&
-  \v{\m w} \equiv \sum_{k\in\T_{LD}\p{M}} w_k \ket{k},
-\end{align}
-and, by slight abuse of notation we identify an equivalence classes
-$k\in\T_{LD}\p{M}$ with a canonical choice of spins
-$\tilde k\in k\subset\C_N\p{M}$ (such that $k_{ap}=\tilde k_{ap}$,
-with $\tilde k_{ap}$ equal to $\tilde k$ with the $a$-th spin replaced
-by $p$); and we define $\ket{j}=0$ for all $j\in\C_N\p{M}$ with
-repeated values.  The matrix $\check{\m h}$ has dimensions
+  \v{\m w} \equiv \sum_{k\in\T_{LD}\p{M}} w_k \ket{k}.
+\end{align}
+In order to make sense of $\check{\m h}$ as written above, for each
+$k\in\T_{LD}\p{M}$ we enforce a canonical choice of spins
+$\tilde k\in k\subset\C_N\p{M}$; we define $k_{ap}$ to be the
+equivalence class of $\tilde k_{ap}$ (where $\tilde k_{ap}$ equal to
+$\tilde k$ with the $a$-th spin $k_a$ replaced by $p$); and we define
+$\ket{k}=0$ if $\tilde k$ has any repeated values.  The matrix
+$\check{\m h}$ has dimensions
 $\abs{\T_{LD}\p{M}}\times\abs{\T_{LD}\p{M}}$, where

### 2020-03-10 collective_sun/sun_perturbation.tex
fix old/incorrect section reference

@@
 ${N \choose 2}\times{N \choose 2}\sim N^2\times N^2$ matrix
-$\check{\m h}_2$.  When both he interaction Hamiltonian $H_0$
-translationally invariant, however, this eigenvalue problem can be
-reduced to that of diagonalizing a matrix of dimensions
-$\sim N\times N$; we perform this reduction in Section
-\ref{sec:exploiting_symmetries}.
+$\check{\m h}_2$.  When the interaction Hamiltonian $H_0$ and
+perturbation $\V_M$ are translationally invariant, however, this
+eigenvalue problem can be reduced to that of diagonalizing a matrix of
+dimensions $\sim N\times N$; we perform this reduction in Section
+\ref{sec:trans_inv}.
 
@@
 \subsection{Translational invariance}
+\label{sec:trans_inv}
 

### 2020-03-10 collective_sun/sun_perturbation.tex
minor changes to text

@@
 \end{align}
-which implies that the generalized eigenvalue problem in
-\eqref{eq:H_VM_psi_eig} can be written as
+which allows us to write the generalized eigenvalue problem in
+\eqref{eq:H_VM_psi_eig} as
 \begin{align}
@@
 $\tilde k\in k\subset\C_N\p{M}$; we define $k_{ap}$ to be the
-equivalence class of $\tilde k_{ap}$ (where $\tilde k_{ap}$ equal to
-$\tilde k$ with the $a$-th spin $k_a$ replaced by $p$); and we define
-$\ket{k}=0$ if $\tilde k$ has any repeated values.  The matrix
+equivalence class of $\tilde k_{ap}$ (where $\tilde k_{ap}$ is equal
+to $\tilde k$ with the $a$-th spin $k_a$ replaced by $p$); and we
+define $\ket{k}=0$ if $\tilde k$ has any repeated values.  The matrix
 $\check{\m h}$ has dimensions
@@
 symmetric states.  The expansion of operators in a basis for this
-joint manifold likewise requires multiplying several operators and
+extended manifold likewise requires multiplying several operators and
 projecting the resulting product onto $\M_0$.  Specifically, if
@@
   &&
-  \C\p{a} \equiv \f{\p{\sum_j a_j}!}{\prod_ja_j!}
-  &&
   \ket{\tilde a} \equiv \bigotimes_{j=1}^n \ket{j}^{\otimes a_j},
+  &&
+  \C\p{a} \equiv \f{\p{\sum_j a_j}!}{\prod_ja_j!},
 \end{align}
@@
 in practice, one can simply sum over $\alpha\in\A_n\p{M}$ and set
-$\beta\equiv b-a+\alpha$.  Note that, by slight abuse of notation, the
+$\beta=b-a+\alpha$.  Note that, by slight abuse of notation, the
 operator $Q$ on the left of \eqref{eq:multi_body_eval} acts on an

### 2020-03-10 collective_sun/sun_perturbation.tex
explicitly exclude the empty set in the domain of diagrams

@@
 where $g_S$ is determined the number of dots at the intersection of
-circles $\set{w_j:j\in S}$.  We denote an assignment of distinct
-indices according to choices of $g_S$ for all $S\in\PP\p{\ZZ_p}$ by
-$g$, where $\PP\p{\ZZ_p}$ is the power set (i.e.~set of all subsets)
-of $\ZZ_p$.  Given a vector $\m M$ of the dimensions $M_j$, we then
-denote the set of all valid index assignments $g$ by $\G\p{\m M}$.
-For consistency, all valid index assignments $g\in\G\p{\m M}$ must
-assign exactly $M_j$ indices to the dimension-$M_j$ tensor $w_j$,
-i.e.~
-\begin{align}
-  \sum_{S\in\PP\p{\ZZ_p}\,:\,j\in S} g_S
-  = \sum_{R\in\PP\p{\ZZ_p\setminus\set{j}}} g_{\set{j}\cup R}
+circles $\set{w_j:j\in S}$.  Denoting the set of all nonempty subsets
+of $\ZZ_p$ by $\PP^*\p{\ZZ_p}$, we denote an assignment of distinct
+indices according to choices of $g_S$ for all $S\in\PP^*\p{\ZZ_p}$ by
+$g$.  Given a vector $\m M$ of the dimensions $M_j$, we then denote
+the set of all valid index assignments $g$ by $\G\p{\m M}$.  For
+consistency, all valid index assignments $g\in\G\p{\m M}$ must assign
+exactly $M_j$ indices to the dimension-$M_j$ tensor $w_j$, i.e.~
+\begin{align}
+  \sum_{S\in\PP^*\p{\ZZ_p}\,:\,j\in S} g_S
+  = \sum_{R\in\PP^*\p{\ZZ_p\setminus\set{j}}} g_{\set{j}\cup R}
   = M_j
 \end{align}
-for all $j\in\ZZ_p$.  In order to make the index assignments
-determined by choices of $g_S$ for all $S\in\PP\p{\ZZ_p}$ unique, we
-also enforce that $g_{\set{}}=0$ for all $g\in\G\p{\m M}$, which is
-equivalent to saying that every dot in the corresponding Venn diagram
-must lie within some circle.  The set of valid index assignments
-$\G\p{\m M}$ is then essentially the set of all Venn diagrams of the
-form in \eqref{eq:venn_diagram} with $p=\dim\p{\m M}$ circles and a
-total of $M_j$ dots within circle $w_j$.
+for all $j\in\ZZ_p$.  The set of valid index assignments $\G\p{\m M}$
+is then essentially the set of all Venn diagrams of the form in
+\eqref{eq:venn_diagram} with $p=\dim\p{\m M}$ circles and a total of
+$M_j$ dots within circle $w_j$.
 
@@
   \C_N\p{g}
-  \equiv \set{ k \in \bigotimes_{S\in\PP\p{\ZZ_p}} \C_N\p{g_S}
+  \equiv \set{ k \in \bigotimes_{S\in\PP^*\p{\ZZ_p}} \C_N\p{g_S}
     : \t{all elements of $k$ are distinct} },
@@
 where each $k\in\C_N\p{g}$ has the natural decomposition
-$k=\p{k_S:S\in\PP\p{\ZZ_p}}$ with $k_S\in\C_N\p{g_S}$.  In words,
+$k=\p{k_S:S\in\PP^*\p{\ZZ_p}}$ with $k_S\in\C_N\p{g_S}$.  In words,
 $\C_N\p{g}$ consists of all the ways to choose, for each subset $S$ of
@@
 \begin{align}
-  k_j \equiv \bigcup_{S\in\PP\p{\ZZ_p} : j\in S} k_S
-  = \bigcup_{R\in\PP\p{\ZZ_p\setminus\set{j}}} k_{\set{j}\cup R},
+  k_j \equiv \bigcup_{S\in\PP^*\p{\ZZ_p} : j\in S} k_S
+  = \bigcup_{R\in\PP^*\p{\ZZ_p\setminus\set{j}}} k_{\set{j}\cup R},
 \end{align}
@@
   \S\p{g} \equiv \sp{\prod_{j\in\ZZ_p} \C\p{ g_S : j\in S }}
-  \sp{\prod_{S\in\PP\p{\ZZ_p}} \p{g_S!}^\abs{S}}
+  \sp{\prod_{S\in\PP^*\p{\ZZ_p}} \p{g_S!}^\abs{S}}
   = \prod_{j\in\ZZ_p} M_j!,
@@
 The simplest diagrams have $g_S$ filled dots (``$\bullet$'') at the
-intersection of the circles $j\in S\in\PP\p{\ZZ_p}$, and are simply
-equal to the associated scalar $w\p{g}$ defined in
-\eqref{eq:diagram_factors}, e.g.
+intersection of the circles $j\in S$, and are simply equal to the
+associated scalar $w\p{g}$ defined in \eqref{eq:diagram_factors}, e.g.
 \begin{align}

### 2020-03-17 collective_sun/sun_perturbation.tex
re-wrote (and corrected) the section on using symmetries to reduce the multi-body eigenvalue problem

@@
 
+% alphanumeric footnotes
+\renewcommand*{\thefootnote}{\alph{footnote}}
+
 %%% text markup
@@
     \p{h_{pk} \bra{\set{k,q}} + h_{qk} \bra{\set{p,k}}}
-    - \tilde h_{pq} \bra{\set{p,q}}}.
+    - \tilde h_{pq} \bra{\set{p,q}}},
+  &&
+  \v{\m w} \equiv \sum_{p<q} w_{pq} \ket{\set{p,q}}.
   \label{eq:V2_eig_mat}
@@
 $h_{kk}=w_{Okk}=0$.  Solving the general eigenvalue problem in
-\eqref{eq:V2_eig} requires diagonalizing the
+\eqref{eq:V2_eig} nominally requires diagonalizing the
 ${N \choose 2}\times{N \choose 2}\sim N^2\times N^2$ matrix
@@
 eigenvalue problem can be reduced to that of diagonalizing a matrix of
-dimensions $\sim N\times N$; we perform this reduction in Section
-\ref{sec:trans_inv}.
+dimensions $\sim N\times N$, and isotropy can reduce this problem even
+further.  We discuss these reductions in Section
+\ref{sec:multi_body_symmetries}.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Constructing multi-body excitation energy eigenstates}
+\section{Constructing excitation energy eigenstates}
 \label{sec:eigenstates}
@@
 permutation of its tensor factors; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is
-a choice of $M$ distinct spins that $O$ acts on; and $\C_N\p{n}$ is
-the set of all distinct choices of $n$ elements from a set of size
+a choice of $M$ distinct spins that $O$ acts on; and $\C_N\p{M}$ is
+the set of all distinct choices of $M$ elements from a set of size
 $N$.
@@
 which implies that $\diag\v h=\sum_dh_{0,d}$ is a scalar.  We thus
-find that,
+find that
 \begin{align}
@@
   = \sum_{k\in\C_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}}
-  \sum_{a=1}^M h_{p k_a} w_k \Pi_{p k_a} O_k \ket\psi.
+  \sum_{a\in\ZZ_M} h_{p k_a} w_k \Pi_{p k_a} O_k \ket\psi.
 \end{align}
@@
 $k$.  Defining
-$k_{ap}\equiv\p{k_1,k_2,\cdots,k_{a-1},p,k_{a+1},\cdots,k_M}$, i.e.~an
-element of $\C_N\p{M}$ in which the $a$-th number $k_a$ has been
-replaced by $p$, we can simplify
+$k_{a:p}\equiv\p{k_1,k_2,\cdots,k_{a-1},p,k_{a+1},\cdots,k_M}$,
+i.e.~$k$ with the $a$-th number $k_a$ replaced by $p$, we can simplify
 \begin{align}
@@
   = \Pi_{p k_a} O_k \Pi_{p k_a} \ket\psi
-  = O_{k_{ap}}\ket\psi,
+  = O_{k_{a:p}}\ket\psi,
 \end{align}
@@
   = \sum_{k\in\C_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}}
-  \sum_{a=1}^M h_{pk_a} w_{k_{ap}} O_k \ket\psi.
+  \sum_{a\in\ZZ_M} h_{pk_a} w_{k_{a:p}} O_k \ket\psi.
 \end{align}
@@
   &&
-  \m h\circ\m w \equiv \sum_{a=1}^M \m h\c_a\m w,
+  \m h\circ\m w \equiv \sum_{a\in\ZZ_M} \m h\c_a\m w,
 \end{align}
@@
   \p{\m h\circ\m w}_k
-  \equiv \sum_{\substack{p\in\ZZ_N\\p\notin k}} \sum_{a=1}^M
-  h_{pk_a} w_{k_{ap}}
+  \equiv \sum_{\substack{p\in\ZZ_N\\p\notin k}} \sum_{a\in\ZZ_M}
+  h_{pk_a} w_{k_{a:p}}
   = \sum_{\substack{p\in\ZZ_N\\p\notin k}}
@@
 \begin{align}
-  \m h\circ\m w - \tilde{\m h} * \m w = \Delta \m w.
+  \m h\circ\m w - \tilde{\m h} * \m w = \Delta \m w,
+\end{align}
+which can be written in the form
+\begin{align}
+  \check{\m h} \c \v{\m w} = \Delta \v{\m w},
   \label{eq:H_VM_psi_eig}
+\end{align}
+where
+\begin{align}
+  \check{\m h} \equiv \sum_{k\in\C_N\p{M}}
+  \sum_{\substack{p\in\ZZ_N\\p\notin k}} \sum_{a\in\ZZ_M}
+  h_{pk_a} \op{k}{k_{a:p}}
+  - \sum_{k\in\C_N\p{M}} \tilde h_k \op{k},
+  &&
+  \v{\m w} \equiv \sum_{k\in\C_N\p{M}} w_k \ket{k}.
 \end{align}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Translational invariance}
-\label{sec:trans_inv}
+\subsection{Translational invariance, isotropy, and other symmetries}
+\label{sec:multi_body_symmetries}
 
@@
 $M$-body perturbations $\V_M$ that generate states of definite
-interaction energy nominally requires diagonalizing a matrix with
-dimensions ${N\choose M}\times{N\choose M}\sim N^M\times N^M$.  Here,
-we simplify this problem to that of diagonalizing a matrix with
-dimensions
-$\sim{N-1\choose M-1}\times{N-1\choose M-1}\sim N^{M-1}\times N^{M-1}$
-when both the interaction Hamiltonian $H_0$ and the perturbation
-$\V_M$ are translationally invariant.  This simplification reduces the
-computational cost of solving the $M$-body eigenvalue problem by a
-factor of $\O\p{N^3}$.
+interaction energy nominally requires diagonalizing the matrix
+$\check{\m h}$ with dimensions
+${N\choose M}\times{N\choose M}\sim N^M\times N^M$.  If the
+interaction Hamiltonian $H_0$ and multi-body perturbation $\V_M$ have
+any symmetries, such as translational invariance isotropy, then we can
+use these symmetry to reduce the eigenvalue problem in
+\eqref{eq:H_VM_psi_eig}.  Translational invariance, for example, can
+reduce this problem to diagonalizing a matrix with dimensions
+$\sim{N-1\choose M-1}\times{N-1\choose M-1}\sim N^{M-1}\times
+N^{M-1}$, and isotropy can further reduce the problem size by factors
+that depend on the lattice geometry.  The complexity of diagonalizing
+a matrix with dimensions $d\times d$ is $\O\p{d^3}$, so even a
+reduction of $d$ by a factor of 2 yields practically significant gains
+(a factor of $2^3=8$) in the runtime of diagonalization.  Here, we
+provide a general prescription for exploiting the symmetries of $H_0$
+and $\V_M$ to simplify the eigenvalue problem in
+\eqref{eq:H_VM_psi_eig}.
+
+Symmetries of $H_0$ and $\V_M$ essentially break the set of choices
+$\C_N\p{M}$ into equivalence classes $\S\p{M}$\footnote{For brevity,
+  we will suppress the explicit dependence of $\S\p{M}$ on $N$, as
+  well as the dependence of $\S\p{M}$ on any other factors such as
+  lattice geometry.}, where elements $s\in\S\p{M}$ are subsets
+$s\subset\C_N\p{M}$ that are invariant under the appropriate symmetry
+group action.  In the case of a translationally invariant and
+isotropic system with two-body perturbations, for example, each
+equivalence class $s\in\S_2$ would consist of all pairs
+$\p{p,q}\in\C_N\p{2}$ that are a fixed distance $\abs{p-q}=d_s$ apart,
+i.e.~$s=\set{\p{p,q}\in\C_N\p{2}:\abs{p-q}=d_s}$.
 
 Our task is essentially to express the multi-body eigenvalue problem
-in \eqref{eq:H_VM_psi_eig} in a basis that obeys both permutational
-symmetry and translational invariance.  To this end, we first expand
-the coupling tensor $\m w$ in a complete basis for symmetric tensors:
-\begin{align}
-  \m w = \sum_{k\in\C_N\p{M}} w_k \m m_k,
-  &&
-  \m m_k \equiv \sum_{\sigma\in P_M} \m e_{\sigma\p{k}},
-  \label{eq:sym_tensor}
-\end{align}
-where $P_M$ is the set of all permutations of $M$ symbols, and
-$\m e_\ell$ is a unit tensor with one in the entry indexed by $\ell$
-and zeros everywhere else.  On a $D$-dimensional lattice of linear
-size $L$, translational invariance implies that the coefficients $w_k$
-remain the same if all spins indexed by $k$ are shifted by the same
-displacement $d\in\ZZ_L^D$.  To account for this invariance, we denote
-a translation of all spins in $k$ by a displacement $d\in\ZZ_L^D$ by
-$k_d$, and define the translation-symmetrized basis tensors
-\begin{align}
-  \m m_k^{\t{TS}}
-  \equiv \f1{\N_k^{\t{TS}}} \sum_{d\in\ZZ_L^D} \m m_{k_d}
-  = \f1{\N_k^{\t{TS}}} \sum_{d\in\ZZ_L^D} \sum_{\sigma\in P_M}
-  \m e_{\sigma(k_d)},
-\end{align}
-where the normalization factor $\N_k^{\t{TS}}$ enforces that the
-entries of $\m m_k^{\t{TS}}$ are all either zero or one, which
-otherwise might not occur if some translations of $k$ results in the
-same choice of spins.  We then define the set $\T_{LD}\p{M}$ of
-equivalence classes in $\C_N\p{M}$ under arbitrary translations on
-$D$-dimensional lattice of linear size $L$, and expand
-\begin{align}
-  \m w = \sum_{k\in\T_{LD}\p{M}} w_k \m m_k^{\t{TS}}.
-  \label{eq:sym_shift_tensor}
-\end{align}
-Having expanded $\m w$ in a basis that respects permutational symmetry
-and translational invariance, we need to express the eigenvalue
-equation \eqref{eq:H_VM_psi_eig} in the same basis.  We therefore
-expand
-\begin{align}
-  \m h\circ\m w
-  = \sum_{k\in\T_{LD}\p{M}} \sum_{p\in\ZZ_L^D} \sum_{a=1}^M
-  h_{pk_a} w_{k_{ap}} \m m_k^{\t{TS}},
-  &&
-  \tilde{\m h} * \m w = \sum_{k\in\T_{LD}\p{M}}
-  \tilde h_k w_k \m m_k^{\t{TS}},
-\end{align}
-which allows us to write the generalized eigenvalue problem in
-\eqref{eq:H_VM_psi_eig} as
-\begin{align}
-  \check{\m h}\c\v{\m w} = \Delta\v{\m w},
+in \eqref{eq:H_VM_psi_eig} in a basis that obeys the symmetries under
+consideration.  To this end, for any equivalence class $s\in\S\p{M}$
+we define a uniform superposition $\ket{s}$ over all choices $k\in s$,
+as well as the projector $\P_\S$ onto the space of uniform
+superpositions:
+\begin{align}
+  \ket{s} \equiv \f1{\sqrt{\abs{s}}} \sum_{k\in s} \ket{k},
+  &&
+  \P_{\S_M} \equiv \sum_{s\in\S\p{M}} \op{s},
+\end{align}
+where $\abs{s}$ is the number of elements in $s$.  We then expand the
+coefficient vector $\v{\m w}$ in this symmetrized basis:
+\begin{align}
+  \v{\m w} = \sum_{k\in\C_N\p{M}} w_k \ket{k}
+  = \sum_{s\in\S_M} w_s \sqrt{\abs{s}} \ket{s},
+\end{align}
+where $w_s\equiv w_k$ for any $k\in s$.  Finally, we define the
+projection of $\check{\m h}$ onto the symmetric subspace:
+\begin{align}
+  \check{\m h}_{\S_M}
+  \equiv \P_{\S_M} \check{\m h} \P_{\S_M}
+  = \sum_{r,s\in\S_M} h_{rs} \op{r}{s},
 \end{align}
@@
 \begin{align}
-  \check{\m h}
-  \equiv \sum_{k\in\T_{LD}\p{M}} \sum_{p\in\ZZ_L^D} \sum_{a=1}^M
-  h_{pk_a} \op{k}{k_{ap}}
-  - \sum_{k\in\T_{LD}\p{M}} \tilde h_k \op{k},
-  &&
-  \v{\m w} \equiv \sum_{k\in\T_{LD}\p{M}} w_k \ket{k}.
-\end{align}
-In order to make sense of $\check{\m h}$ as written above, for each
-$k\in\T_{LD}\p{M}$ we enforce a canonical choice of spins
-$\tilde k\in k\subset\C_N\p{M}$; we define $k_{ap}$ to be the
-equivalence class of $\tilde k_{ap}$ (where $\tilde k_{ap}$ is equal
-to $\tilde k$ with the $a$-th spin $k_a$ replaced by $p$); and we
-define $\ket{k}=0$ if $\tilde k$ has any repeated values.  The matrix
-$\check{\m h}$ has dimensions
-$\abs{\T_{LD}\p{M}}\times\abs{\T_{LD}\p{M}}$, where
-$\abs{\T_{LD}\p{M}} \approx { N-1 \choose M-1 }/M \sim N^{M-1}$.
+  h_{rs} \equiv \sum_{k,\ell\in\C_N} h_{k\ell} \bk{r|k} \bk{\ell|s}
+  = \f1{\sqrt{\abs{r}\abs{s}}}
+  \sum_{\substack{k\in r\\\ell\in s}} h_{k\ell}.
+\end{align}
+In practice, it can be computationally cheaper to construct
+\begin{align}
+  \check{\m h}_{\S_M} = \sum_{s\in\S_M} \ket{s} \f1{\sqrt{\abs{s}}}
+  \sum_{k\in s} \sum_{\substack{p\in\ZZ_N\\p\notin k}} \sum_{a\in\ZZ_M}
+  \f{h_{p k_a}}{\sqrt{\abs{\sp{k_{a:p}}}}} \bra{\sp{k_{a:p}}},
+\end{align}
+where $k_a$ is the $a$-th element of $k$;
+$k_{a:p}\equiv\p{k_1,k_2,\cdots,k_{a-1},p,k_{a+1},\cdots,k_M}$ is $k$,
+but with the $a$-th number $k_a$ replaced by $p$; and $\sp{k_{a:p}}$
+is the equivalence class of $k_{a:p}$.  The reduced multi-body
+eigenvalue problem is then
+\begin{align}
+  \check{\m h}_{\S_M} \c \v{\m w} = \Delta \v{\m w}.
+\end{align}
 

### 2020-03-17 collective_sun/sun_perturbation.tex
changed a section title and a bit of text

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Operator products restricted to the permutationally symmetric
-  manifold}
+\section{Projecting products of operators onto the permutationally
+  symmetric manifold}
 \label{sec:sym_prod}
@@
 systematic way.  To this end, we derive a general diagrammatic
-expansion of these products, as well as rules to simplify the
-resulting expansions so as to minimize their computational cost.
-These diagrammatic methods allow us to algorithmically evaluate
-products such as \eqref{eq:example_prod} from the bare multi-body
-operator data $\p{w^{(j)},O},\p{v,Q},\p{w^{(k)},O}$.  We then work out
-some special cases of products that are relevant to the perturbation
-theory in Sections \ref{sec:single_body_pert} and
-\ref{sec:two_body_pert}.
+expansion for these products, as well as a set of rules to simplify
+these expansions so as to minimize their computational cost.  These
+diagrammatic methods allow us to algorithmically evaluate products
+such as \eqref{eq:example_prod} from the bare multi-body operator data
+$\p{w^{(j)},O},\p{v,Q},\p{w^{(k)},O}$.  We then work out some special
+cases of products that are relevant to the perturbation theory in
+Sections \ref{sec:single_body_pert} and \ref{sec:two_body_pert}.
 

### 2020-03-17 collective_sun/sun_perturbation.tex
changed expression and fixed error in section in symmetries in the multi-body eigenvalue problem

@@
   \equiv \P_{\S_M} \check{\m h} \P_{\S_M}
-  = \sum_{r,s\in\S_M} h_{rs} \op{r}{s},
-\end{align}
-where
-\begin{align}
-  h_{rs} \equiv \sum_{k,\ell\in\C_N} h_{k\ell} \bk{r|k} \bk{\ell|s}
-  = \f1{\sqrt{\abs{r}\abs{s}}}
-  \sum_{\substack{k\in r\\\ell\in s}} h_{k\ell}.
-\end{align}
-In practice, it can be computationally cheaper to construct
-\begin{align}
-  \check{\m h}_{\S_M} = \sum_{s\in\S_M} \ket{s} \f1{\sqrt{\abs{s}}}
+  = \sum_{s\in\S_M} \ket{s} \sp{
     \sum_{k\in s} \sum_{\substack{p\in\ZZ_N\\p\notin k}} \sum_{a\in\ZZ_M}
-  \f{h_{p k_a}}{\sqrt{\abs{\sp{k_{a:p}}}}} \bra{\sp{k_{a:p}}},
-\end{align}
-where $k_a$ is the $a$-th element of $k$;
+    \f{h_{p k_a}}{\sqrt{\abs{s}\abs{\sp{k_{a:p}}}}} \bra{\sp{k_{a:p}}}
+    - \tilde h_s \bra{s}},
+\end{align}
+where $\tilde h_s\equiv \tilde h_k$ for any $k\in s$; $k_a$ is the
+$a$-th element of $k$;
 $k_{a:p}\equiv\p{k_1,k_2,\cdots,k_{a-1},p,k_{a+1},\cdots,k_M}$ is $k$,

### 2020-03-17 collective_sun/sun_perturbation.tex
added comment about reducing the multi-body eigenvalue problem when H_0 has more symmetries than V_M

@@
 \end{align}
+In general, symmetries of the interaction Hamiltonian $H_0$ give the
+matrix $\check{\m h}$ a block diagonal structure, with blocks of size
+$\abs{\S_M}\times\abs{\S_M}$ corresponding to different orbits under
+the symmetry group action.  These blocks can be classified, for
+example, by a choice of wavenumber
+$k_s\in\ZZ_{\abs{s}}\times 2\pi/\abs{s}$ for each equivalence class
+$s\in\S_M$.  Even if the multi-body operator $\V_M$ does not obey the
+same symmetries as $H_0$, this block diagonal structure can be used to
+reduce the full multi-body eigenvalue problem to that of diagonalizing
+each block.  We defer this reduction to future work.
 

### 2020-03-17 collective_sun/sun_perturbation.tex
fixed minor notational inconsistency

--- 2020-03-17 collective_sun/sun_perturbation.tex
@@
 Symmetries of $H_0$ and $\V_M$ essentially break the set of choices
-$\C_N\p{M}$ into equivalence classes $\S\p{M}$\footnote{For brevity,
-  we will suppress the explicit dependence of $\S\p{M}$ on $N$, as
-  well as the dependence of $\S\p{M}$ on any other factors such as
-  lattice geometry.}, where elements $s\in\S\p{M}$ are subsets
+$\C_N\p{M}$ into equivalence classes $\S_M$\footnote{For brevity, we
+  will suppress the explicit dependence of $\S_M$ on $N$, as well as
+  the dependence of $\S_M$ on any other factors such as lattice
+  geometry.}, where elements $s\in\S_M$ are subsets
 $s\subset\C_N\p{M}$ that are invariant under the appropriate symmetry
@@
 in \eqref{eq:H_VM_psi_eig} in a basis that obeys the symmetries under
-consideration.  To this end, for any equivalence class $s\in\S\p{M}$
-we define a uniform superposition $\ket{s}$ over all choices $k\in s$,
-as well as the projector $\P_\S$ onto the space of uniform
+consideration.  To this end, for any equivalence class $s\in\S_M$ we
+define a uniform superposition $\ket{s}$ over all choices $k\in s$, as
+well as the projector $\P_\S$ onto the space of uniform
 superpositions:
@@
   &&
-  \P_{\S_M} \equiv \sum_{s\in\S\p{M}} \op{s},
+  \P_{\S_M} \equiv \sum_{s\in\S_M} \op{s},
 \end{align}

### 2020-03-19 collective_sun/sun_perturbation.tex
overhaul of main text and notation in the multi-body notes these notes are now structures to consider general M-body perturbations from the start

@@
 \newcommand{\G}{\mathcal{G}}
-\renewcommand{\H}{\mathcal{H}}
 \newcommand{\I}{\mathcal{I}}
-\renewcommand{\L}{\mathcal{L}}
 \newcommand{\M}{\mathcal{M}}
 \newcommand{\N}{\mathcal{N}}
-\renewcommand{\O}{\mathcal{O}}
 \renewcommand{\P}{\mathcal{P}}
-\newcommand{\Q}{\mathcal{Q}}
 \renewcommand{\S}{\mathcal{S}}
 \newcommand{\T}{\mathcal{T}}
-\newcommand{\V}{\mathcal{V}}
-\newcommand{\X}{\mathcal{X}}
-
-\newcommand{\DD}{\mathbb{D}}
+\newcommand{\Z}{\mathcal{Z}}
+
 \newcommand{\PP}{\mathbb{P}}
-\newcommand{\RR}{\mathbb{R}}
 \renewcommand{\SS}{\mathbb{S}}
@@
 
+\let\var\relax
+\DeclareMathOperator{\var}{var}
 \DeclareMathOperator{\diag}{diag}
@@
 
-\title{Perturbing SU($n$)-symmetric interactions}%
+\title{SU($n$) spin models near the Heisenberg critical point}%
 \author{Michael A. Perlin}%
@@
 
+\red{TODO: write an introduction}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{The SU($n$) Heisenberg model}
+
 We consider an array of $N$ multilevel spins with SU($n$)-symmetric
@@
 \begin{align}
-  H_0 = \sum_{p<q} h_{pq} \Pi_{pq},
-  &&
-  \Pi_{pq} \equiv \sum_{\mu,\nu} s_{\mu\nu}^{(p)} s_{\nu\mu}^{(q)},
+  H_0 = \sum_{\p{p,q}\in\C_N\p{2}} h_{pq} \Pi_{pq},
+  &&
+  \Pi_{pq} \equiv \sum_{\mu,\nu\in\ZZ_n} s_{\mu\nu}^{(p)} s_{\nu\mu}^{(q)},
   \label{eq:H_0}
 \end{align}
-where $p,q$ index individual spins; $\mu,\nu$ index states in an
-orthonormal basis for the $n$-dimensional Hilbert space $\H_n$ of a
-single spin; $h_{pq}$ are scalar coefficients; the operator
+where $p,q$ index individual spins; $\C_N\p{M}$ is the set of all
+choices of $M$ spins out of $N$; $h_{pq}$ are scalar coupling
+constants; $\mu,\nu$ index states in an orthonormal basis for the
+$n$-dimensional Hilbert space of a single spin; $\ZZ_N$ is the set of
+integers modulo $N$; the operator
 $s_{\mu\nu}^{(p)}\equiv\op{\mu}{\nu}_p$ flips the state of spin $p$ to
 $\ket\mu$ from $\ket\nu$; and the operator $\Pi_{pq}$ permutes the
-states of spins $p$ and $q$.  The permutation operator $\Pi_{pq}$ is a
-multilevel generalization of the SU(2)-symmetric spin interaction
-$\v s_p\c\v s_q$ for $\v s=\p{s_\x,s_\y,s_\z}$, with
-$s_\alpha=\sigma_\alpha/2$ and $\sigma_\alpha$ a Pauli matrix for an
-individual spin.
-
-We wish to determine the effective dynamics induced on the
-ground-state manifold $\M_0$ of $H_0$ by weak single- and two-body
-perturbations of the form
-\begin{align}
-  \V_1 \equiv \sum_X \sum_p v_{Xp} X_p = \sum_X \v v_X\c\v X,
-  &&
-  \V_2 \equiv \sum_O \sum_{p<q} w_{Opq} O_{pq}
-  = \sum_O \v{\m w}_O\c \v{\m O}
-  \label{eq:perturbations}
-\end{align}
-where $p,q$ index individual spins; $v_{Xp}$ and $w_{Opq}$ are real
-numbers (coefficients); $X$ is a trace-zero single-spin operator on
-$\H_n$; $O$ is a trace-zero two-spin operator on $\H_n\otimes\H_n$
-that obeys pair-wise permutational symmetry,
-i.e.~$O_{pq} = \Pi_{pq} O_{pq} \Pi_{pq} = O_{qp}$; $\v v_X$ and $\v X$
-are $N$-component vectors defined by
-\begin{align}
-  \v v_X \equiv \sum_p v_{Xp} \ket{p},
-  &&
-  \v X \equiv \sum_p X_p \ket{p};
-\end{align}
-and finally $\v{\m w_O}$ and $\v{\m O}$ are ${N \choose 2}$-component
-vectors defined by
-\begin{align}
-  \v{\m w}_O \equiv \sum_{p<q} w_{Opq} \ket{\set{p,q}},
-  &&
-  \v{\m O} \equiv \sum_{p<q} O_{pq} \ket{\set{p,q}},
-\end{align}
-where the basis vectors $\ket{\set{p,q}}$ are labeled by a choice of
-two distinct spins $p,q$.
+states of spins $p$ and $q$.  The permutation operator $\Pi_{pq}$ is
+essentially a multilevel generalization of the SU(2)-symmetric spin
+interaction $\v s_p\c\v s_q$, with $\v s\equiv\p{\v s_\x,s_\y,s_\z}$,
+$s_\alpha\equiv\sigma_\alpha/2$, and $\sigma_\alpha$ a single-spin
+Pauli operator.
 
@@
 all negative, then the ground-state manifold $\M_0$ of the interaction
-Hamiltonian $H_0$ consists of permutationally symmetric states that
-are simultaneous $+1$ eigenstates of all permutation operators
-$\Pi_{pq}$.  We will adopt the restriction that all $h_{pq}<0$
-throughout these notes, but note that this restriction can be relaxed
-to the assumption that the initial state of all spins is
-permutationally symmetric (e.g.~a spin-polarized state).  We will also
-assume throughout these notes that the permutationally symmetric
-manifold $\M_0$ is gapped away from all other states by an interaction
-energy difference that is large compared to any coupling between
-$\M_0$ and its orthogonal complement.  Power-law couplings of the form
-$h_{pq}=-h/\abs{p-q}^\alpha$ on a $D$-dimensional lattice, for
-example, always yield a non-vanishing spectral gap when $\alpha\le D$
-(see Section \ref{sec:spectral_gap}).
-
-The effective Hamiltonian $H_M$ induced on the ground-state manifold
-$\M_0$ by an $M$-body perturbation $\V_M$ through second order in
-perturbation theory is given by\cite{bravyi2011schrieffer,
-  perlin2019effective}
-\begin{align}
-  H_M = H_M^{(1)} + H_M^{(2)},
-  &&
-  H_M^{(1)} = \P_0 \V_M \P_0,
-  &&
-  H_M^{(2)} = - \P_0 \V_M \E \V_M \P_0,
-  &&
-  \E \equiv \sum_{\Delta\ne0} \f{\P_\Delta}{\Delta},
-  \label{eq:H_M}
-\end{align}
-where $\P_\Delta$ is a projector onto the eigenspace of the
-interaction Hamiltonian $H_0$ with interaction energy $\Delta$ above
-that of permutationally symmetric manifold $\M_0$; that is, the
-interaction energy of states in the support of $\P_\Delta$ is
-$E_0+\Delta$, where
-\begin{align}
-  E_0 = \sum_{p<q} h_{pq}
-\end{align}
-is the interaction energy of permutationally symmetric states
-$\ket\psi\in\M_0$.
+Hamiltonian $H_0$ consists of permutationally symmetric (PS) states
+that are simultaneous $+1$ eigenstates of all permutation operators
+$\Pi_{pq}$.  We will assume that all $h_{pq}<0$ throughout these
+notes, but comment that this restriction can be relaxed to the
+assumption that the initial state of all spins is permutationally
+symmetric (e.g.~a spin-polarized state).  We will also assume that the
+PS manifold $\M_0$ is gapped away from all orthogonal states by an
+interaction energy $\Delta_{\t{gap}}>0$.  Power-law couplings of the
+form $h_{pq}\sim-1/\abs{p-q}^\alpha$ on a $D$-dimensional lattice, for
+example, always yield a non-vanishing spectral gap $\Delta_{\t{gap}}$
+for long-range interactions with $\alpha\le D$ (see Appendix
+\ref{sec:spectral_gap}).
+
+The PS manifold $\M_0$ is spanned by a basis of states
+$\ket{m}=\ket{m_1,m_2,\cdots,m_n}$ with a definite occupation number
+$m_\mu$ of the single-spin state $\mu$, and $\sum_\mu m_\mu=N$.  In
+the case of $n=2$, for example, the PS manifold $\M_0$ is precisely
+the Dicke manifold\cite{dicke1954coherence} of $N$ qubits, spanned by
+Dicke states $\set{\ket{m_0,m_1}}$ for which $m_0+m_1=N$ and the
+integer $m_0$ ($m_1$) indicates the total occupation number of qubit
+state $\ket{0}$ ($\ket{1}$).  The dimension of the PS manifold $\M_0$
+is determined by the number of ways to assign $N$ spins to $n$ states,
+which is ${N+n-1 \choose n-1}\sim N^{n-1}$.  The energy of any PS
+state $\ket\psi\in\M_0$ with respect to the interaction Hamiltonian
+$H_0$ in \eqref{eq:H_0} is
+\begin{align}
+  E_0 = \sum_{\p{p,q}\in\C_N\p{2}} h_{pq}.
+\end{align}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Single-body perturbations}
-\label{sec:single_body_pert}
-
-To compute the effective Hamiltonian induced by the single-body
-perturbation $\V_1$, we use the permutational symmetry of the
-ground-state manifold to simplify
-\begin{align}
-  H_1^{(1)} = \sum_{X,p} v_{Xp} \P_0 X_p \P_0
-  \EQPS \sum_{X,p} v_{Xp} X_1
-  \EQPS \sum_X \bar v_X \col{X},
-  \label{eq:H_1_1}
-\end{align}
-where $\EQPS$ denotes equality under a restriction to the
-permutationally symmetric manifold $\M_0$, and
-\begin{align}
-  \bar v_X \equiv \f1N \sum_p v_{Xp},
-  &&
-  \col{X} \equiv \sum_p X_p,
-\end{align}
-are respectively the mean value in $\v v_X$ and the collective version
-of $X$.  In order to compute the second order effective Hamiltonian
-$H_1^{(2)}$, we first choose an arbitrary permutationally symmetric
-state $\ket\psi\in\M_0$ and expand (see Section \ref{sec:H_V1_psi}):
-\begin{align}
-  H_0 \V_1 \ket\psi
-  = E_0 \V_1 \ket\psi
-  + \sum_X \sum_p \p{\sum_q h_{pq} v_{Xq} - h_p v_{Xp}} X_p
-  \ket\psi,
-  &&
-  h_p \equiv \sum_q h_{pq},
-  \label{eq:H_V1_psi}
-\end{align}
-where we define $h_{pq}$ for all $p,q$ (as opposed to only $p<q$) by
-enforcing $h_{pq}=h_{qp}$ and $h_{pp}=0$.  We thus find that a
-perturbation $\V_1^\Delta\ket\psi$ (indexed by $\Delta$) is an
-eigenvector of the interaction Hamiltonian $H_0$ if its coefficient
-vectors $\v v_X^\Delta$ all satisfy the eigenvalue equation
-\begin{align}
-  \p{\m h - \diag\v h}\c\v v = \Delta \v v,
-  \label{eq:V1_eig}
-\end{align}
-where $\m h\equiv\sum_{p,q}h_{pq}\op{p}{q}$ is a matrix of all
-couplings $h_{pq}$; the vector $\v h\equiv\sum_{p,q}h_{pq}\ket{p}$ is
-the sum of all columns of $\m h$; and $\diag\v h$ is a matrix with
-$\v h$ on the diagonal and zeroes everywhere else.  Solving the
-general eigenvalue problem in \eqref{eq:V1_eig} requires diagonalizing
-the $N\times N$ matrix $\m h-\diag\v h$.  When the interaction
-Hamiltonian $H_0$ is translationally invariant, however, this
-eigenvalue problem can be solved analytically; we provide the
-corresponding solution at the end of Section \ref{sec:H_V1_psi}.
-
-If we {\it construct} a perturbation $\V_1^\Delta$ with coefficients
-$\v v_X^\Delta$ that satisfy \eqref{eq:V1_eig} with eigenvalue
-$\Delta$, then
-\begin{align}
-  H_0 \V_1^\Delta \ket\psi = \p{E_0 + \Delta} \V_1^\Delta \ket\psi.
-\end{align}
-Interestingly, the energy of the state $\V_1^\Delta\ket\psi$ depends
-only on the coefficients $\v v_X^\Delta$, and is entirely independent
-of the state $\ket\psi$ or the choice of a trace-zero single-spin
-operator $X$ used to build $\V_1^\Delta$.  Finding operators
-$\V_1^\Delta$ that generate eigenvectors of the interaction
-Hamiltonian $H_0$ when they act on permutationally symmetric states
-$\ket\psi\in\M_0$ thus reduces to finding eigenvectors of the
-interaction matrix $\m h-\diag\v h$.
-
-Any coefficient vector $\v v_X$ can be expanded into its projections
-$\v v_X^\Delta$ onto the eigenspace of vectors satisfying
-\eqref{eq:V1_eig} with eigenvalue $\Delta$,
-i.e.~$\v v_X = \sum_\Delta \v v_X^\Delta$, which also allows us to
-expand
-\begin{align}
-  \V_1 = \sum_\Delta \V_1^\Delta,
-  &&
-  \V_1^\Delta \equiv \sum_X \v v_X^\Delta \c \v X,
-\end{align}
-where each operator $\V_1^\Delta$ generates a state with definite
-interaction energy $\Delta$ above that of the permutationally
-symmetric manifold $\M_0$.  We can therefore simplify
-\begin{align}
-  H_1^{(2)}
-  = - \P_0 \V_1 \E \V_1 \P_0
-  = - \sum_{\Delta\ne0} \f1{\Delta}
-  \P_0 \V_1^\Delta \P_\Delta \V_1^\Delta \P_0
-  = - \sum_{\Delta\ne0} \f1{\Delta} \P_0 \p{\V_1^\Delta}^2 \P_0,
-\end{align}
-where the product $\P_0 \p{\V_1^\Delta}^2 \P_0$ is worked out in
-Section \ref{sec:PXYP}.  In total, we find that
-\begin{align}
-  H_1^{(2)}
-  \EQPS \f1{N\p{N-1}} \sum_{X,Y} \sum_{\Delta\ne0}
-  \f{\v v_X^\Delta\c\v v_Y^\Delta}{\Delta}
-  \p{\col{X}\,\col{Y} - N \col{XY}},
-  &&
-  \col{Z} \equiv \sum_p Z_p.
-  \label{eq:H_1_2}
-\end{align}
-where $\EQPS$ denotes equality under a restriction to the
-permutationally symmetric manifold $\M_0$; and $\v v_X^\Delta$ is the
-projection of $\v v_X$ onto the $\Delta$-eigenspace of the matrix
-$\m h-\diag\v h$.
+\section{Breaking SU($n$) symmetry (perturbatively)}
+\label{eq:pert_theory}
+
+In order to induce non-trivial dynamics for initially PS states
+$\ket\psi\in\M_0$, which are otherwise stationary eigenstates of the
+interaction Hamiltonian $H_0$ in \eqref{eq:H_0}, we will add to the
+Hamiltonian an operator of the general form \red{[TODO: consider more
+  general sums of such operators]}
+\begin{align}
+  X_{\m w} \equiv \sum_{k\in\C_N\p{M}} w_k X_k,
+  \label{eq:pert}
+\end{align}
+where $X$ is an $M$-spin operator that is invariant under arbitrary
+permutation of its tensor factors; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is
+a choice of $M$ distinct spins that $X_k$ acts on; $\C_N\p{M}$ is the
+set of all choices of $M$ elements from a set of size $N$; and $\m w$
+is a dimension-$M$ tensor with scalar entries $w_k$.  Even though only
+${N\choose M}$ entries of $\m w$ are used to define $X_{\m w}$ in
+\eqref{eq:pert}, for later convenience we define $w_k$ for all indices
+$k\in\ZZ_N^M$ by enforcing that $\m w$ is symmetric with respect to
+any permutation of its indices, and that $w_k=0$ for any $k$ with
+repeated values. \red{[TODO: relax the restrictions on $X$ and $\m w$
+  to allow for more general perturbations, e.g.~of the form
+  $s_\x\otimes s_\y-s_\y\otimes s_\x$]}
+
+Single-body ($M=1$) operators correspond to classical external fields,
+e.g.~$s_\z$ for a magnetic field, and two-body ($M=2$) operators
+correspond to two-body interactions that deviate from $H_0$, such as
+$s_\z\otimes s_\z$ interactions in an XXZ model\red{[TODO:cite]} near
+the Heisenberg critical point.  While we would be hard-pressed to find
+a physical implementation of multi-body spin interactions with $M>2$,
+considering this case will be necessary to construct low-lying
+eigenstates of $H_0$.  Constructing low-lying eigenstates of $H_0$
+will in turn enable simulating the dynamics induced by
+non-perturbative additions $X_{\m w}$ to the SU($n$) Heisenberg model,
+with norms $\norm{X_{\m w}}\sim\Delta_{\t{gap}}$.
+
+When $X_{\m w}$ is sufficiently small, we can treat its effect on the
+PS manifold $\M_0$ perturbatively.  The effective Hamiltonian
+$H_{\t{eff}}$ induced by $X_{\m w}$ on the ground-state manifold
+$\M_0$ through second order in perturbation theory
+is\cite{bravyi2011schrieffer, perlin2019effective}
+\begin{align}
+  H_{\t{eff}} = H_{\t{eff}}^{(1)} + H_{\t{eff}}^{(2)},
+  &&
+  H_{\t{eff}}^{(1)} = \P_0 X_{\m w} \P_0,
+  &&
+  H_{\t{eff}}^{(2)} = - \sum_{\Delta\ne0}
+  \f1\Delta \P_0 X_{\m w} \P_\Delta X_{\m w} \P_0,
+  \label{eq:H_eff}
+\end{align}
+where $\P_\Delta$ is a projector onto the eigenspace $\E_\Delta$ of
+the interaction Hamiltonian $H_0$ with interaction energy $\Delta$
+above that of PS manifold $\M_0$.  In particular, $\P_0$ is a
+projector onto $\M_0$ itself, and $\E_0=\M_0$.  The first order
+effective Hamiltonian $H_{\t{eff}}^{(1)}$ is straightforward to
+simplify using the permutational symmetry of $\M_0$:
+\begin{align}
+  H_{\t{eff}}^{(1)} \EQPS \bar w\,\col{X},
+  \label{eq:H_eff_1}
+\end{align}
+where $\EQPS$ denotes equality under a restriction to $\M_0$; $\bar w$
+is an average of $w_k$ over spin choices $k\in\C_N\p{M}$; and
+$\col{X}$ is the collective version of $X$, i.e.
+\begin{align}
+  \bar w \equiv { N \choose M }^{-1} \sum_{k\in\C_N\p{M}} w_k,
+  &&
+  \col{X} \equiv \sum_{k\in\C_N\p{M}} X_k,
+\end{align}
+An immediate consequence of \eqref{eq:H_eff_1} is the fact that
+perturbing SU(2)-symmetric interactions by two-body $s_\z\otimes s_\z$
+interactions yields the one-axis twisting (OAT)
+Hamiltonian\cite{kitagawa1993squeezed, ma2011quantum}
+$H_{\t{OAT}}=\chi S_\z^2$ at first order in perturbation theory, with
+the OAT strength $\chi$ equal to the mean coefficient of the two-body
+$s_\z\otimes s_\z$ terms.
+
+The second order effective Hamiltonian $H_{\t{eff}}^{(2)}$ is not as
+straightforward to simplify due to the presence of a projector
+$\P_\Delta$ onto the manifold $\E_\Delta$ of states with excitation
+energy $\Delta$ (with respect to $H_0$).  This projector essentially
+picks off the part of $X_{\m w}$ that is strictly off-diagonal with
+respect to the ground- and excited-state manifolds $\E_0$ and
+$\E_\Delta$.  To decompose $X_{\m w}$ into components that generate
+states of definite excitation energy when acting on PS states $\M_0$,
+we pick an arbitrary state $\ket\psi\in\M_0$ and expand (see Appendix
+\ref{sec:eigenstates})
+\begin{align}
+  H_0 X_{\m w} \ket\psi
+  = E_0 X_{\m w} \ket\psi + X_{\hat{\m h}\p{\m w}} \ket\psi,
+  \label{eq:diagnosis}
+\end{align}
+where $\hat{\m h}$ is a linear map on the space of dimension-$M$
+tensors, determined entirely by the couplings $h_{pq}$ of the
+interaction Hamiltonian $H_0$.  The image $\hat{\m h}\p{\m w}$ of
+$\m w$ under $\hat{\m h}$ is then itself a dimension-$M$ tensor that
+defines the $M$-body operator $X_{\hat{\m h}\p{\m w}}$.  To write out
+$\hat{\m h}\p{\m w}$ explicitly, we first define a matrix
+$\m h\equiv\sum_{p,q\in\ZZ_N} h_{pq} \op{p}{q}$ of all couplings in
+$H_0$ by enforcing that $h_{pq}=h_{qp}$ and $h_{pp}=0$.  We denote a
+contraction of $\m h$ with the $a$-th index of $\m w$ by
+$\m h \circ_a\m w$, such that the $k$-th entry of $\m h \circ_a\m w$
+is
+\begin{align}
+  \p{\m h \circ_a \m w}_k
+  = \sum_{p\in\ZZ_N\setminus k} h_{k_a p} w_{k_{a:p}},
+\end{align}
+where $k\equiv\p{k_1,k_2,\cdots,k_M}$ is a list of indices;
+$k_{a:p}\equiv\p{k_1,k_2,\cdots,k_{a-1},p,k_{a+1},\cdots,k_M}$ is the
+same list $k$, but with the $a$-th index $k_a$ replaced by $p$; and we
+explicitly exclude terms with $p\in k$ that vanish due to the fact
+that $\m h$ and $\m w$ are strictly off-diagonal.  We then define a
+dimension-$M$ tensor $\tilde{\m h}$ with entries
+\begin{align}
+  \tilde h_k \equiv \sum_{\substack{p\in k\\q\in\ZZ_N\setminus k}} h_{pq}
+\end{align}
+for all $k\in\ZZ_N^M$, and we denote an element-wise (Kronecker)
+product of $\tilde{\m h}$ and $\m w$ by $\tilde{\m h} * \m w$, with
+entries $\tilde h_k w_k$.  These definitions allow us to expand
+\begin{align}
+  \hat{\m h}\p{\m w}
+  \equiv \sum_{a\in\ZZ_M} \m h \circ_a \m w - \tilde{\m h} * \m w.
+\end{align}
+Viewed an operator on the vector space of dimension-$M$ tensors, the
+map $\hat{\m h}$ is real and symmetric under transposition, and is
+therefore diagonalizable.  Diagonalizing this operator yields
+eigenvalues $\Delta$ and eigenvectors $\m m_\Delta$ for which
+$\check{\m h}\p{\m m_\Delta}=\Delta\m m_\Delta$.  Finding eigenvalues
+and eigenvectors of $\hat{\m h}$ nominally requires diagonalizing a
+matrix of dimensions
+${N\choose M}\times{N\choose M}\sim N^M\times N^M$, but symmetries of
+$\hat{\m h}$ such as translational invariance or isotropy can
+drastically reduce the complexity of this eigenvalue problem.  We
+discuss this reduction in Appendix \ref{sec:symmetries}.
+
+Defining $\m w_\Delta$ as the projection of the tensor $\m w$ onto the
+$\Delta$-eigenspace of $\check{\m h}$, we can expand
+\begin{align}
+  \m w = \sum_\Delta \m w_\Delta,
+  &&
+  \hat{\m h}\p{\m w} = \sum_\Delta \Delta \m w_\Delta,
+\end{align}
+and in turn
+\begin{align}
+  X_{\m w} = \sum_\Delta X_{\m w_\Delta},
+  &&
+  X_{\hat{\m h}\p{\m w}} = \sum_\Delta \Delta X_{\m w_\Delta}.
+\end{align}
+Substituting these expansions into \eqref{eq:diagnosis}, we find
+\begin{align}
+  H_0 X_{\m w} \ket\psi
+  = \sum_\Delta\p{E_0 + \Delta} X_{\m w_\Delta} \ket\psi,
+\end{align}
+which implies that the action of $X_{\m w_\Delta}$ on a PS state
+$\ket\psi\in\M_0$ generates a state of definite interaction energy
+$E_0+\Delta$.  It follows that
+$\P_\Delta X_{\m w} \P_0 = \P_\Delta X_{\m w_\Delta} \P_0 = X_{\m
+  w_\Delta} \P_0$, and so
+\begin{align}
+  H_{\t{eff}}^{(2)} = -\sum_{\Delta\ne0}
+  \f1\Delta \P_0 X_{\m w_\Delta} X_{\m w_\Delta} \P_0.
+\end{align}
+The product $\P_0 X_{\m w_\Delta} X_{\m w_\Delta} \P_0$ generally
+includes $r$-body operators for all positive integers $r\le2M$.  We
+provide a general prescription for simplifying such products in
+Appendix \ref{sec:operator_product}, and explicitly work out the
+single-body ($M=1$) and two-body ($M=2$) cases in Appendices
+\ref{sec:single_pair_prod} and \ref{sec:two_pair_prod}.  If $X$ is a
+single-body operator ($M=1$), then
+\begin{align}
+  H_{\t{eff}}^{(2)}
+  = \sum_{\Delta\ne0} \f{\var\p{\m w_\Delta}}{\p{N-1}\Delta}
+  \p{\col{X}^2 - N \col{X^2}},
+  &&
+  \var\p{\m m} \equiv \f1N \sum_{p\in\ZZ_N} \p{m_p-\bar m}^2,
+\end{align}
+which in the SU(2) case implies that an inhomogeneous magnetic field
+($X\propto\sigma_\z$) yields an OAT Hamiltonian at second order
+perturbation theory.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Two-body perturbations}
-\label{sec:two_body_pert}
-
-The first-order effective Hamiltonian $H_2^{(1)}$ induced on the
-permutationally symmetric manifold $\M_0$ by the two-body perturbation
-$\V_2$ in \eqref{eq:perturbations} is simply the restriction of $\V_2$
-onto $\M_0$:
-\begin{align}
-  H_2^{(1)} = \sum_O \sum_{p<q} w_{Opq} \P_0 O_{pq} \P_0
-  \EQPS \sum_O \sum_{p<q} w_{Opq} O_{1,2}
-  \EQPS \sum_O \bar w_O \col{O},
-\end{align}
-where
-\begin{align}
-  \bar w_O \equiv {N \choose 2}^{-1} \sum_{p<q} w_{Opq},
-  &&
-  \col{O} \equiv \sum_{p<q} O_{pq},
-\end{align}
-are an average of $w_{Opq}$ and a collective version of the two-spin
-operator $O_{pq}$.  If the two-spin operator
-$O=X\otimes Y+Y\otimes X$, then
-\begin{align}
-  H_2^{(1)} \EQPS \f12 \sum_{\p{X,Y}} \bar w_{\p{X,Y}}
-  \p{\col{X}\,\col{Y} + \col{Y}\,\col{X} - \col{XY} - \col{YX}}.
-\end{align}
-In order to compute the second-order effective Hamiltonian
-$H_2^{(2)}$, as before we pick an arbitrary permutationally symmetric
-state $\ket\psi\in\M_0$ and expand (see Section \ref{sec:eigenstates})
-\begin{align}
-  H_0 \V_2 \ket\psi
-  = E_0 \V_2 \ket\psi
-  + \sum_O \sum_{p<q}
-  \sp{\sum_k \p{h_{pk} w_{Okq} + h_{qk} w_{Opk}}
-    - \tilde h_{pq} w_{Opq}}
-  O_{pq} \ket\psi,
-  \label{eq:H_V2_psi}
-\end{align}
-where to keep notation compact, we define
-\begin{align}
-  \tilde h_{pq} \equiv h_p + h_q - 2 h_{pq},
-  &&
-  h_p \equiv \sum_q h_{pq},
-\end{align}
-and we define $w_{Opq}$ for all $p,q$ (as opposed to only $p<q$) by
-enforcing $w_{Opq}=w_{Oqp}$ and $w_{Opp}=0$.  We thus find that the
-vector $\V_2^\Delta\ket\psi$ (indexed by $\Delta$) is an eigenvector
-of the interaction Hamiltonian $H_0$ if its coefficients
-$\v{\m w}_O^\Delta$ all satisfy the eigenvalue equation
-\begin{align}
-  \check{\m h}_2 \c \v{\m w} = \Delta \v{\m w},
-  \label{eq:V2_eig}
-\end{align}
-where
-\begin{align}
-  \check{\m h}_2
-  \equiv \sum_{p<q} \ket{\set{p,q}} \sp{\sum_{k\notin\set{p,q}}
-    \p{h_{pk} \bra{\set{k,q}} + h_{qk} \bra{\set{p,k}}}
-    - \tilde h_{pq} \bra{\set{p,q}}},
-  &&
-  \v{\m w} \equiv \sum_{p<q} w_{pq} \ket{\set{p,q}}.
-  \label{eq:V2_eig_mat}
-\end{align}
-We explicitly exclude $k\in\set{p,q}$ in \eqref{eq:V2_eig_mat} because
-these terms vanish in \eqref{eq:H_V2_psi} from the fact that
-$h_{kk}=w_{Okk}=0$.  Solving the general eigenvalue problem in
-\eqref{eq:V2_eig} nominally requires diagonalizing the
-${N \choose 2}\times{N \choose 2}\sim N^2\times N^2$ matrix
-$\check{\m h}_2$.  When the interaction Hamiltonian $H_0$ and
-perturbation $\V_M$ are translationally invariant, however, this
-eigenvalue problem can be reduced to that of diagonalizing a matrix of
-dimensions $\sim N\times N$, and isotropy can reduce this problem even
-further.  We discuss these reductions in Section
-\ref{sec:multi_body_symmetries}.
-
-If we {\it construct} a perturbation $\V_2^\Delta$ with coefficients
-$\v{\m w}_O^\Delta$ that satisfy \eqref{eq:V2_eig} with eigenvalue
-$\Delta$, then
-\begin{align}
-  H_0 \V_2^\Delta \ket\psi = \p{E_0 + \Delta} \V_2^\Delta \ket\psi,
-\end{align}
-which implies that $\V_2^\Delta\ket\psi$ is a state of definite energy
-that depends only on the coefficients $\m w_O^\Delta$.  Similarly to
-the case of single-body perturbations, we can decompose any vector
-$\v{\m w}_O$ into its projections $\v{\m w}_O^\Delta$ onto the
-eigenspace of vectors satisfying \eqref{eq:V2_eig} with eigenvalue
-$\Delta$, i.e.~$\v{\m w}_O=\sum_\Delta\v{\m w}_O^\Delta$, which allows
-us to expand
-\begin{align}
-  \V_2 = \sum_\Delta \V_2^\Delta,
-  &&
-  \V_2^\Delta \equiv \sum_O \v{\m w}_O^\Delta \c \v{\m O},
-\end{align}
-and in turn write
-\begin{align}
-  H_2^{(2)} = - \P_0 \V_2 \E \V_2 \P_0
-  = -\sum_{\Delta\ne0} \f1\Delta \P_0 \p{\V_2^\Delta}^2 \P_0,
-\end{align}
-where the product $\P_0 \p{\V_2^\Delta}^2 \P_0$ is simplified in
-Section \ref{sec:POQP}.  In the special case that
-\begin{align}
-  \V_2 = \sum_{p<q} w_{pq} s_\z^{(p)} s_\z^{(q)},
-\end{align}
-where $s_\z\equiv\p{\op\up-\op\dn}/2$ is an SU(2) spin-$z$ operator,
-the second-order effective Hamiltonian $H_2^{(2)}$ takes the form
-\begin{align}
-  H_2^{(2)} \simeq_\PS W_4 S_\z^4 - W_2 S_\z^2,
-\end{align}
-where $\simeq_\PS$ denotes equality up to scalar terms and a
-restriction to the permutationally symmetric manifold;
-$S_\z\equiv\sum_p s_\z^{(p)}$ is a collective SU(2) spin operator; and
-$W_k$ are scalar coefficients.  In order to expand the coefficients
-$W_k$, we define
-\begin{align}
-  \v{\m w}=\sum_{p<q} w_{pq} \ket{\set{p,q}},
-  &&
-  \v w \equiv \sum_{p,q} w_{pq} \ket{p},
-\end{align}
-where $w_{pq}$ is defined for all $p,q$ (as opposed to only $p<q$) by
-enforcing $w_{pq}=w_{qp}$ and $w_{pp}=0$.  The vector $\v w$ is the
-sum of all columns of the matrix
-$\m w\equiv\sum_{p,q}w_{pq}\op{p}{q}$.  We then denote the projection
-of $\v{\m w}$ onto the $\Delta$-eigenspace of the matrix
-$\check{\m h}_2$ defined in \eqref{eq:V2_eig_mat} by
-$\v{\m w}_\Delta$, and define
-\begin{align}
-  w_{\t{columns}}^{(2)} \equiv \sum_{\Delta\ne0}
-  \f{\v w_\Delta\c\v w_\Delta}{\Delta},
-  &&
-  w_{\t{pairs}}^{(2)} \equiv \sum_{\Delta\ne0}
-  \f{\v{\m w}_\Delta\c\v{\m w}_\Delta}{\Delta},
-\end{align}
-in terms of which
-\begin{align}
-  W_4 \equiv \f{w_{\t{columns}}^{(2)} - w_{\t{pairs}}^{(2)}}{N!_4},
-  &&
-  W_2 \equiv \f{w_{\t{columns}}^{(2)} - 2 w_{\t{pairs}}^{(2)}}
-  {4N!_2} + \f12 \p{3N-4} W_4,
-\end{align}
-where
-\begin{align}
-  N!_k \equiv \prod_{j=0}^{k-1} \p{N-j} = \f{N!}{\p{N-k}!}
-\end{align}
-denotes a falling factorial.
+\section{Beyond ground-state perturbation theory}
+\label{eq:excited_states}
+
+In order to find the effective Hamiltonian $H_{\t{eff}}^{(2)}$ induced
+by the $M$-body operator $X_{\m w}$ on the PS manifold $\M_0$ at
+second order in perturbation theory, we constructed states with
+definite excitation energy $\Delta$ with respect to $H_0$.  These
+states are generated by the action of an $M$-body operator
+$X_{\m w_\Delta}$ on a PS state $\ket\psi\in\M_0$.  We denote the
+manifold of all states that can be generated by the action of some
+$M$-body operator, but not any $r$-body operator with $r<M$, by
+$\M_M$; we will refer to $\M_M$ as the $M$-body excitation manifold.
+We then denote the a direct sum of all $\M_r$ for $r\le M$ by
+$\ul{\M}_M$, i.e.~$\ul{\M}_M\equiv\bigoplus_{r\le M}\M_r$.  The
+manifold $\ul{\M}_M$ consists of all states accessible by the action
+of $r$-body operators with $r\le M$ on PS states in $\M_0$.
+
+By construction, the $M$-body operator $X_{\m w}$ generally couples PS
+states in $\M_0$ to asymmetric states in $\ul{\M}_M\setminus\M_0$.  If
+the coupling between $\M_0$ and its complement is weak compared to the
+spectral gap $\Delta_{\t{gap}}$ of $H_0$, then we can treat the effect
+of $X_{\m w}$ perturbatively and derive an effective Hamiltonian
+$H_{\t{eff}}$ on $\M_0$.  This ground-state perturbation theory
+formally breaks down when the operator norm
+$\norm{X_{\m w}}\ge\Delta_{\t{gap}}/2$, in which case the operator
+$X_{\m w}$ first couples $\M_0$ to $\ul{\M}_M$, then $\ul{\M}_M$ to
+$\ul{\M}_{2M}$, and so on.  Even if
+$\norm{X_{\m w}}>\Delta_{\t{gap}}/2$, however, the energetic penalty
+incurred by $H_0$ from breaking the permutational symmetry of an
+initial state $\ket\psi\in\M_0$ suppresses the leakage of population
+into manifolds $\M_r$ of increasing $r$.  One might therefore capture
+the dynamics of an initial PS state $\ket\psi\in\M_0$ by projecting
+$X_{\m w}$ onto $\ul{\M}_{r_{\t{max}}}$ for some $r_{\t{max}}\ge M$,
+resulting in a theory analogous to first-order perturbation theory on
+$\ul{\M}_{r_{\t{max}}}$.  In the perturbative regime
+$\norm{X_{\m w}}<\Delta_{\t{gap}}/2$, such a theory captures all
+virtual processes through order $2\floor{r_{\t{max}}/M}+1$ in a formal
+perturbation theory on $\M_0$.  In the non-perturbative regime
+$\norm{X_{\m w}}\ge\Delta_{\t{gap}}/2$, the error incurred by
+truncating the Hilbert space of a spin system past
+$\ul{\M}_{r_{\t{max}}}$ is not clear a priori.  Nonetheless,
+simulations restricted to $\ul{\M}_{r_{\t{max}}}$ can be benchmarked
+by examining the maximum populations within all $\M_r$ for
+$r\le r_{\t{max}}$ throughout simulation.  As long as these
+populations fall off with increasing $r$ and the maximal population
+within large-$r$ manifolds $\M_r$ is small, one should expect such
+simulations to faithfully capture (at least qualitatively) the
+dynamical behavior of the spin system under consideration.
 
@@
 $N=L^D$ spins, the spectral gap of the interaction Hamiltonian is (see
-Section \ref{sec:H_V1_psi})
+Appendix \ref{sec:single_body_eigenstates})
 \begin{align}
@@
 denoting an association that ignores the cyclic structure of $\ZZ_L$;
-$k_{\t{SE}}\in\ZZ_L\times2\pi/L$ is a wavenumber for a singly-excited
-spin-wave state; and the norm $\abs{d}$ for a vector $d$ on a periodic
-lattice is implicitly understood to mean the smallest Euclidean
-distance of $d$ from a fixed origin.  In order to yield the smallest
-excitation energy $\Delta$, the wavenumber $k_{\t{SE}}$ should
-maximize the contribution of the cosine term above, which is achieved
-by a wavenumber that minimizes the oscillations of this term when
-integrated over the entire lattice.  A suitable candidate for a
-minimal wavenumber is $k_{\t{SE}}=\p{2\pi/L,0,0,\cdots}$, which
-corresponds to an excitation energy
+$k_{\t{SE}}\in\ZZ_L^D\times2\pi/L$ is a wavenumber for a
+singly-excited spin-wave state; and the norm $\abs{d}$ for a vector
+$d$ on a periodic lattice is implicitly understood to mean the
+smallest Euclidean distance of $d$ from a fixed origin.  In order to
+yield the smallest excitation energy $\Delta$, the wavenumber
+$k_{\t{SE}}$ should maximize the contribution of the cosine term
+above, which is achieved by a wavenumber that minimizes the
+oscillations of this term when integrated over the entire lattice.  A
+suitable candidate for a minimal wavenumber is
+$k_{\t{SE}}=\p{2\pi/L,0,0,\cdots}$, which corresponds to an excitation
+energy
 \begin{align}
@@
 \end{align}
-As $\epsilon\to0$ the discrete sum over $x$ becomes an integral that
-avoids an infinitesimal region about the origin, i.e.
+As $\epsilon\to0$ the discrete sum over $x$ is well approximated by an
+integral that avoids an infinitesimal region about the origin, i.e.
 \begin{align}
@@
 In this section, we simplify the products of the form
-$H_0\V_M\ket\psi$ to arrive at expansions such as \eqref{eq:H_V1_psi}
-and \eqref{eq:H_V2_psi}, which define conditions under which a
-multi-body perturbation $\V_M$ generates eigenstates of the
+$H_0X_{\m w}\ket\psi$ to arrive at the expansion in
+\eqref{eq:diagnosis}, which defines conditions under which a
+multi-body perturbation $X_{\m w_\Delta}$ generates eigenstates of the
 SU($n$)-symmetric interaction Hamiltonian $H_0$ in \eqref{eq:H_0} when
 acting on a permutationally symmetric state $\ket\psi\in\M_0$.  Here
-$\V_M$ is an $M$-body perturbation of the form
-\begin{align}
-  \V_M \equiv \sum_{k\in\C_N\p{M}} w_k O_k.
-\end{align}
-where $O$ is an $M$-spin operator that is invariant under arbitrary
+$X_{\m w}$ is an $M$-body perturbation defined by
+\begin{align}
+  X_{\m w} \equiv \sum_{k\in\C_N\p{M}} w_k X_k,
+  \label{eq:pert_copy}
+\end{align}
+where $X$ is an $M$-spin operator that is invariant under arbitrary
 permutation of its tensor factors; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is
-a choice of $M$ distinct spins that $O$ acts on; and $\C_N\p{M}$ is
-the set of all distinct choices of $M$ elements from a set of size
-$N$.
+a choice of $M$ distinct spins that $X_k$ acts on; $\C_N\p{M}$ is the
+set of all choices of $M$ elements from a set of size $N$; and $\m w$
+is a dimension-$M$ tensor with scalar entries $w_k$.  Even though only
+${N\choose M}$ entries of $\m w$ are used to define $X_{\m w}$ in
+\eqref{eq:pert_copy}, for later convenience we define $w_k$ for all
+indices $k\in\ZZ_N^M$ by enforcing that $\m w$ is symmetric with
+respect to any permutation of its indices, and that $w_k=0$ for any
+$k$ with repeated values. \red{[TODO: relax the restrictions on $X$
+  and $\m w$ to allow for more general perturbations, e.g.~of the form
+  $s_\x\otimes s_\y-s_\y\otimes s_\x$]}
 
@@
 \subsection{Single-body excitations}
-\label{sec:H_V1_psi}
-
-As a toy calculation, we first consider the simple case of a
-single-body perturbation and simplify
-\begin{align}
-  H_0 \V_1 \ket\psi
-  = \sum_k \sum_{p<q} h_{pq} v_k \Pi_{pq} X_k \ket\psi,
-  \label{eq:H_V1_psi_start}
-\end{align}
-where without loss of generality we consider a single-body
-perturbation $\V_1$ built out of only one single-spin operator $X$.
-The sum in \eqref{eq:H_V1_psi_start} has terms with $k\in\set{p,q}$,
-and terms with $k\notin\set{p,q}$.  In the case of $k\notin\set{p,q}$,
-the permutation operator $\Pi_{pq}$ commutes with $X_k$ and
-annihilates on $\ket\psi$, leaving terms at each fixed $k$ of the form
-\begin{align}
-  \sum_{\substack{p<q\\k\notin\set{p,q}}} h_{pq}
-  = \sum_{p<q} h_{pq} - \sum_{\substack{p<q\\k\in\set{p,q}}} h_{pq}
-  = E_0 - \f12 \sum_{\substack{p,q\\k\in\set{p,q}}} h_{pq}
+\label{sec:single_body_eigenstates}
+
+In order to build familiarity with the problem at hand, we first
+consider the simple case of a single-body perturbation and simplify
+\begin{align}
+  H_0 X_{\m w} \ket\psi
+  = \sum_{\substack{\p{p,q}\in\C_N\p{2}\\k\in\ZZ_N}}
+  h_{pq} w_k \Pi_{pq} X_k \ket\psi.
+  \label{eq:single_body_diagnosis_start}
+\end{align}
+The sum in \eqref{eq:single_body_diagnosis_start} has terms with
+$k\in\set{p,q}$, and terms with $k\notin\set{p,q}$.  In the case of
+$k\notin\set{p,q}$, the permutation operator $\Pi_{pq}$ commutes with
+$X_k$ and annihilates on $\ket\psi$, leaving terms at each fixed $k$
+of the form
+\begin{align}
+  \sum_{\substack{\p{p,q}\in\C_N\p{2}\\k\notin\set{p,q}}} h_{pq}
+  = \sum_{\p{p,q}\in\C_N\p{2}} h_{pq}
+  - \sum_{\substack{\p{p,q}\in\C_N\p{2}\\k\in\set{p,q}}} h_{pq}
+  = E_0 - \f12 \sum_{\substack{p,q\in\ZZ_N\\k\in\set{p,q}}} h_{pq}
   = E_0 - h_k,
 \end{align}
-where we define $h_{pq}$ for all $p,q$ (as opposed to only $p<q$) by
-enforcing $h_{pq}=h_{qp}$ and $h_{pp}=0$; and
-$h_k \equiv \sum_\ell h_{k\ell}$.  The terms in
-\eqref{eq:H_V1_psi_start} with $k\notin\set{p,q}$ are then
-\begin{align}
-  \sum_{\substack{p<q\\k\notin\set{p,q}}}
+where we define $h_{pq}$ for arbitrary $p,q$ by enforcing
+$h_{pq}=h_{qp}$ and $h_{pp}=0$; and $h_k \equiv \sum_\ell h_{k\ell}$.
+The terms in \eqref{eq:single_body_diagnosis_start} with
+$k\notin\set{p,q}$ are then
+\begin{align}
+  \sum_{\substack{\p{p,q}\in\C_N\p{2}\\k\notin\set{p,q}}}
+  h_{pq} w_k \Pi_{pq} X_k \ket\psi
+  = \sum_{k\in\ZZ_N} \p{E_0 - h_k} w_k X_k \ket\psi
+  = E_0 X_{\m w} \ket\psi - \sum_{k\in\ZZ_N} h_k w_k X_k \ket\psi
+\end{align}
+while the terms in \eqref{eq:single_body_diagnosis_start} with
+$k\in\set{p,q}$ take the form
+\begin{align}
+  \sum_{\substack{\p{p,q}\in\C_N\p{2}\\k\in\set{p,q}}}
   h_{pq} v_k \Pi_{pq} X_k \ket\psi
-  = \sum_k \p{E_0 - h_k} v_k X_k \ket\psi
-  = E_0 \V_1 \ket\psi - \sum_k h_k v_k X_k \ket\psi
-\end{align}
-while the terms in \eqref{eq:H_V1_psi_start} with $k\in\set{p,q}$ take
-the form
-\begin{align}
-  \sum_{\substack{p<q\\k\in\set{p,q}}}
-  h_{pq} v_k \Pi_{pq} X_k \ket\psi
-  = \sum_{k<q} h_{kq} v_k X_q \ket\psi
-  + \sum_{p<k} h_{pk} v_k X_p \ket\psi
-  = \sum_{k,q} h_{kq} v_k X_q \ket\psi
+  = \sum_{\p{k,q}\in\C_N\p{2}} h_{kq} w_k X_q \ket\psi
+  + \sum_{\p{p,k}\in\C_N\p{2}} h_{pk} w_k X_p \ket\psi
+  = \sum_{k,q\in\ZZ_N} h_{kq} w_k X_q \ket\psi
 \end{align}
@@
 \begin{align}
-  H_0 \V_1 \ket\psi
-  = E_0 \V_1 \ket\psi
-  + \sum_p \p{\sum_{q} h_{pq} v_q - h_p v_p}
+  H_0 X_{\m w} \ket\psi
+  = E_0 X_{\m w} \ket\psi
+  + \sum_{p\in\ZZ_N} \p{\sum_{q\in\ZZ_N} h_{pq} w_q - h_p w_p}
   X_p \ket\psi.
-  \label{eq:H_V1_psi_end}
-\end{align}
-The action of a single-body perturbation $\V_1=\sum_p v_p X_p$ on a
-fully symmetric state therefore generates an eigenstate of $H_0$ with
-energy $E_0+\Delta$ if the vector $\v v\equiv\sum_p v_p\ket{p}$
-satisfies the eigenvalue equation
-\begin{align}
-  \p{\m h - \diag\v h} \c \v v = \Delta \v v,
-  \label{eq:H_V1_psi_eig}
-\end{align}
-where $\m h\equiv\sum_{p,q}h_{pq}\op{p}{q}$ is a matrix of all
-couplings $h_{pq}$; the vector $\v h\equiv\sum_{p,q}h_{pq}\ket{p}$ is
-the sum of all columns of $\m h$; and $\diag\v h$ is a matrix with
-$\v h$ on the diagonal and zeroes everywhere else.
+  \label{eq:single_body_diagnosis_end}
+\end{align}
+The action of the single-body perturbation $X_{\m w}$ on a
+permutationally symmetric state therefore generates an eigenstate of
+$H_0$ with energy $E_0+\Delta$ if the vector
+$\v{\m w}\equiv\sum_{k\in\ZZ_N} w_k \ket{k}$ satisfies the eigenvalue
+equation
+\begin{align}
+  \p{\m h - \diag\v h} \c \v{\m w} = \Delta \v{\m w},
+  \label{eq:single_body_eig}
+\end{align}
+where $\m h\equiv\sum_{p,q\in\ZZ_N} h_{pq} \op{p}{q}$ is a matrix of
+all couplings $h_{pq}$; the vector
+$\v h\equiv\sum_{p,q\in\ZZ_N} h_{pq} \ket{p}$ is the sum of all
+columns of $\m h$; and $\diag\v h$ is a matrix with $\v h$ on the
+diagonal and zeroes everywhere else.
 
 When the interaction Hamiltonian $H_0$ is translationally invariant,
-the single-body eigenvalue problem in \eqref{eq:H_V1_psi_eig} is
-exactly solvable.  In this case, the couplings $h_{pq}$ depend only on
-the separation $\abs{p-q}$, so the eigenvectors of $\m h$ are plane
-waves of the form
-\begin{align}
-  \v v_k \equiv \sum_p e^{ip\c k} \ket{p},
+the single-body eigenvalue problem in \eqref{eq:single_body_eig} is
+solvable analytically.  In this case, the couplings $h_{pq}$ depend
+only on the separation $\abs{p-q}$, so the eigenvectors of $\m h$ are
+plane waves of the form
+\begin{align}
+  \v{\m w}_k \equiv \sum_{p\in\ZZ_L^D} e^{ip\c k} \ket{p},
 \end{align}
 where on a $D$-dimensional periodic lattice of $N=L^D$ spins, lattice
-sites are indexed by vectors
-$p\in\ZZ_L^D\simeq\set{0,1,\cdots,L-1}^{\otimes D}$, and wavenumbers
-take on values $k\in\ZZ_L^D\times2\pi/L$.  The corresponding
-eigenvalues of $\m h$ can be determined by expanding
-\begin{align}
-  \m h\c\v v_k = \sum_{p,q} h_{pq} e^{iq\c k} \ket{p}
-  = \sum_{p,d} h_{p,p+d} e^{i\p{p+d}\c k} \ket{p}
-  = \sum_d h_{0,d} e^{id\c k} \v v_k
-  = \sum_d h_{0,d} \cos\p{d\c k} \v v_k,
+sites are indexed by vectors $p\in\ZZ_L^D$, and wavenumbers take on
+values $k\in\ZZ_L^D\times2\pi/L$.  The corresponding eigenvalues of
+$\m h$ can be determined by expanding
+\begin{align}
+  \m h\c{\m w}_k
+  = \sum_{p,q\in\ZZ_L^D} h_{pq} e^{iq\c k} \ket{p}
+  = \sum_{p,d\in\ZZ_L^D} h_{p,p+d} e^{i\p{p+d}\c k} \ket{p}
+  = \sum_{d\in\ZZ_L^D} h_{0,d} e^{id\c k} \v{\m w}_k
+  = \sum_{d\in\ZZ_L^D} h_{0,d} \cos\p{d\c k} \v{\m w}_k,
 \end{align}
 where the imaginary terms vanish because $h_{0,d}=h_{0,-d}$.  The
-remainder of \eqref{eq:H_V1_psi_eig} that we need to sort out is
-$\diag\v h$, where all $h_p=\sum_qh_{pq}=\sum_dh_{0,d}$ are equal,
-which implies that $\diag\v h=\sum_dh_{0,d}$ is a scalar.  We thus
-find that
-\begin{align}
-  \p{\m h - \diag\v h}\c\v v_k = \Delta_k \v v_k,
-  &&
-  \Delta_k \equiv \sum_d h_{0,d} \sp{\cos\p{d\c k}-1}.
+remainder of \eqref{eq:single_body_eig} that we need to sort out is
+$\diag\v h$, where all
+$h_p=\sum_{q\in\ZZ_L^D}h_{pq}=\sum_{d\in\ZZ_L^D}h_{0,d}$ are equal,
+which implies that $\diag\v h=\sum_{d\in\ZZ_L^D}h_{0,d}$ is a scalar.
+We thus find that
+\begin{align}
+  \p{\m h - \diag\v h}\c\v v_k = \Delta_k \v{\m w}_k,
+  &&
+  \Delta_k \equiv \sum_{d\in\ZZ_L^D} h_{0,d} \sp{\cos\p{d\c k}-1}.
 \end{align}
@@
 \subsection{Multi-body excitations}
-\label{sec:eigenstates}
-
-We now consider the full $M$-body product
-\begin{align}
-  H_0 \V_M \ket\psi
-  = \sum_{\substack{\p{p,q}\in\C_N\p{2}\\k\in\C_N\p{M}}} h_{pq} w_k
-  \Pi_{pq} O_k \ket\psi,
-  \label{eq:H_VM_psi_start}
-\end{align}
-where without loss of generality we consider an $M$-body perturbation
-$\V_M$ built out of only one $M$-spin operator $O$.  The sum in
-\eqref{eq:H_VM_psi_start} has terms with $p,q\in k$, terms with
-$p,q\notin k$, and terms with only one of $p$ or $q\in k$.  The
-permutation operator $\Pi_{pq}$ acts trivially on $O_k\ket\psi$ when
+\label{sec:multi_body_eigenstates}
+
+We now consider the full $M$-body case
+\begin{align}
+  H_0 X_{\m w} \ket\psi
+  = \sum_{\substack{\p{p,q}\in\C_N\p{2}\\k\in\C_N\p{M}}}
+  h_{pq} w_k \Pi_{pq} X_k \ket\psi.
+  \label{eq:diagnosis_start}
+\end{align}
+The sum in \eqref{eq:diagnosis_start} has terms with $p,q\in k$, terms
+with $p,q\notin k$, and terms with only one of $p$ or $q\in k$.  The
+permutation operator $\Pi_{pq}$ acts trivially on $X_k\ket\psi$ when
 $p,q\in k$ or $p,q\notin k$, leaving terms at each fixed $k$ of the
@@
   = \sum_{\p{p,q}\in\C_N\p{2}} h_{pq}
-  - \sum_{p\in k} \sum_{\substack{q\in\ZZ_N\\q\notin k}} h_{pq}
-  = E_0 - \sum_{p\in k}
-  \p{\sum_{q\in\ZZ_N} h_{pq} - \sum_{q\in k} h_{pq}},
-\end{align}
-where we first split a sum over $\p{p,q}\in\C_N\p{2}$ with the
-restriction that $p,q\in k~\t{or}~p,q\notin k$ into an unrestricted
-sum minus a remainder, and then similarly split the sum over all spins
-$q$ with the restriction that $q\notin k$ into a sum over all $q$
-minus the remainder $q\in k$.  We thus find that
-\begin{align}
-  \sum_{\substack{\p{p,q}\in\C_N\p{2}\\p,q\in k~\t{or}~p,q\notin k}} h_{pq}
+  - \sum_{\substack{p\in k\\q\notin k}} h_{pq}
   = E_0 - \tilde h_k,
   &&
-  \tilde h_k \equiv \sum_{p\in k}\p{h_p - \sum_{q\in k}h_{pq}},
-  &&
-  h_p \equiv \sum_q h_{pq},
-\end{align}
-which implies that the terms in \eqref{eq:H_VM_psi_start} with
-$p,q\in k$ or $p,q\notin k$ are
-\begin{align}
-  \sum_{k\in\C_N\p{M}} \p{E_0 - \tilde h_k} w_k O_k \ket\psi
-  = E_0 \V_M \ket\psi
-  - \sum_{k\in\C_N\p{M}} \p{\tilde{\m h} * \m w}_k O_k \ket\psi,
+  \tilde h_k \equiv \sum_{\substack{p\in k\\q\in\ZZ_N\setminus k}} h_{pq},
+\end{align}
+where we split a sum over $\p{p,q}\in\C_N\p{2}$ with the restriction
+that $p,q\in k$ or $p,q\notin k$ into an unrestricted sum minus a
+remainder with the complementary restriction.  The terms in
+\eqref{eq:diagnosis_start} with $p,q\in k$ or $p,q\notin k$ are then
+\begin{align}
+  \sum_{k\in\C_N\p{M}} \p{E_0 - \tilde h_k} w_k X_k \ket\psi
+  = E_0 X_{\m w} \ket\psi
+  - \sum_{k\in\C_N\p{M}} \p{\tilde{\m h} * \m w}_k X_k \ket\psi,
 \end{align}
 where $\tilde{\m h}$ is a tensor with entries $\tilde h_k$; and
-$\m x*\m y$ denotes an element-wise (Kronecker) product of $\m x$ and
-$\m y$, with entries $x_ky_k$.  The remaining terms in
-\eqref{eq:H_VM_psi_start} with only one of $p\in k$ or $q\in k$ are
+$\tilde{\m h} * \m w$ denotes an element-wise (Kronecker) product of
+$\tilde{\m h}$ and $\m w$, with entries $\tilde h_k w_k$.  The
+remaining terms in \eqref{eq:diagnosis_start} with only one of
+$p\in k$ or $q\in k$ are
 \begin{align}
   \sum_{k\in\C_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}} \sum_{q\in k}
-  h_{pq} w_k \Pi_{pq} O_k \ket\psi
+  h_{pq} w_k \Pi_{pq} X_k \ket\psi
   = \sum_{k\in\C_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}}
-  \sum_{a\in\ZZ_M} h_{p k_a} w_k \Pi_{p k_a} O_k \ket\psi.
-\end{align}
-where $k\equiv\p{k_1,k_2,\cdots,k_M}$, with $k_a$ the $a$-th entry in
-$k$.  Defining
-$k_{a:p}\equiv\p{k_1,k_2,\cdots,k_{a-1},p,k_{a+1},\cdots,k_M}$,
-i.e.~$k$ with the $a$-th number $k_a$ replaced by $p$, we can simplify
-\begin{align}
-  \Pi_{p k_a} O_k \ket\psi
-  = \Pi_{p k_a} O_k \Pi_{p k_a} \ket\psi
-  = O_{k_{a:p}}\ket\psi,
+  \sum_{a\in\ZZ_M} h_{p k_a} w_k \Pi_{p k_a} X_k \ket\psi.
+\end{align}
+where $k=\p{k_1,k_2,\cdots,k_M}$, with $k_a$ the $a$-th entry in $k$.
+Defining $k_{a:p}$ to be a list that is equal to $k$ but with the
+$a$-th number $k_a$ replaced by $p$,
+$k_{a:p}\equiv\p{k_1,k_2,\cdots,k_{a-1},p,k_{a+1},\cdots,k_M}$, we can
+simplify
+\begin{align}
+  \Pi_{p k_a} X_k \ket\psi
+  = \Pi_{p k_a} X_k \Pi_{p k_a} \ket\psi
+  = X_{k_{a:p}}\ket\psi,
 \end{align}
@@
 \begin{align}
-  \sum_{k\in\C_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}} \sum_{q\in k}
-  h_{pq} w_k  \Pi_{pq} O_k \ket\psi
+  \sum_{k\in\C_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}}
+  \sum_{q\in k} h_{pq} w_k  \Pi_{pq} X_k \ket\psi
   = \sum_{k\in\C_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}}
-  \sum_{a\in\ZZ_M} h_{pk_a} w_{k_{a:p}} O_k \ket\psi.
+  \sum_{a\in\ZZ_M} h_{pk_a} w_{k_{a:p}} X_k \ket\psi.
 \end{align}
@@
 \begin{align}
-  \sum_{k\in\C_N\p{M}} \p{\m h\circ \m w}_k O_k \ket\psi,
-  &&
-  \m h\circ\m w \equiv \sum_{a\in\ZZ_M} \m h\c_a\m w,
+  \sum_{k\in\C_N\p{M}} \sum_{a\in\ZZ_M}
+  \p{\m h \circ_a \m w}_k X_k \ket\psi,
 \end{align}
@@
 \begin{align}
-  \p{\m h\circ\m w}_k
-  \equiv \sum_{\substack{p\in\ZZ_N\\p\notin k}} \sum_{a\in\ZZ_M}
-  h_{pk_a} w_{k_{a:p}}
+  \p{\m h \circ_a \m w}_k
+  \equiv \sum_{\substack{p\in\ZZ_N\\p\notin k}} h_{pk_a} w_{k_{a:p}}
   = \sum_{\substack{p\in\ZZ_N\\p\notin k}}
-  \p{h_{k_1 p} w_{pk_2k_3\cdots k_M}
-    + h_{k_2 p} w_{k_1pk_3\cdots k_M}
-    + \cdots + h_{k_M p} w_{k_1k_2\cdots k_{M-1} p}}.
+  h_{pk_a} w_{k_1 k_2 \cdots k_{a-1} p k_{a+1} \cdots k_M}.
 \end{align}
@@
 \begin{align}
-  H_0 \V_M \ket\psi
-  = E_0 \V_M \ket\psi + \sum_{k\in\C_N\p{M}}
-  \p{\m h\circ\m w - \tilde{\m h} * \m w}_k O_k \ket\psi.
-  \label{eq:H_VM_psi_end}
-\end{align}
-The action of the $M$-body perturbation $\V_M$ on a fully symmetric
-state therefore generates an eigenstate of $H_0$ with energy
-$E_0+\Delta$ if the coefficient tensor $\m w$ satisfies the
-generalized eigenvalue equation
-\begin{align}
-  \m h\circ\m w - \tilde{\m h} * \m w = \Delta \m w,
-\end{align}
-which can be written in the form
-\begin{align}
-  \check{\m h} \c \v{\m w} = \Delta \v{\m w},
-  \label{eq:H_VM_psi_eig}
+  H_0 X_{\m w} \ket\psi
+  = E_0 X_{\m w} \ket\psi + \sum_{k\in\C_N\p{M}}
+  \hat{\m h}\p{\m w}_k X_k \ket\psi,
+  &&
+  \hat{\m h}\p{\m w}
+  \equiv \p{\sum_a \m h \circ_a \m w - \tilde{\m h} * \m w}.
+  \label{eq:diagnosis_end}
+\end{align}
+The action of the $M$-body perturbation $X_{\m w}$ on a
+permutationally symmetric state therefore generates an eigenstate of
+$H_0$ with energy $E_0+\Delta$ if the coefficient tensor $\m w$
+satisfies the generalized eigenvalue equation
+$\hat{\m h}\p{\m w} = \Delta \m w$, which can be written in the form
+\begin{align}
+  \hat{\m h} \c \v{\m w} = \Delta \v{\m w},
+  \label{eq:multi_body_eig}
 \end{align}
@@
 \begin{align}
-  \check{\m h} \equiv \sum_{k\in\C_N\p{M}}
-  \sum_{\substack{p\in\ZZ_N\\p\notin k}} \sum_{a\in\ZZ_M}
-  h_{pk_a} \op{k}{k_{a:p}}
+  \hat{\m h} \equiv \sum_{k\in\C_N\p{M}}
+  \sum_{\substack{p\in\ZZ_N\\p\notin k}}
+  \sum_{a\in\ZZ_M} h_{pk_a} \op{k}{k_{a:p}}
   - \sum_{k\in\C_N\p{M}} \tilde h_k \op{k},
@@
 \end{align}
-If $M=1$, then $\p{\m h\circ\m w}_k=\sum_\ell h_{k\ell}w_\ell$ and
-$\tilde h_k=h_k$, so
-\begin{align}
-  H_0 \V_1 \ket\psi
-  = E_0 \V_1 \ket\psi + \sum_{k\in\ZZ_N}
-  \p{\sum_{\ell\in\ZZ_N} h_{k\ell} w_\ell - h_k w_k} O_k \ket\psi,
-\end{align}
-which is precisely what we found in \eqref{eq:H_V1_psi_end}.  If
-$M=2$, then
-$\p{\m h\circ\m w}_{pq}=\sum_k\p{h_{pk}w_{kq}+h_{qk}w_{pk}}$, so
-\begin{align}
-  H_0 \V_2 \ket\psi
-  = E_0 \V_2 \ket\psi + \sum_{\p{p,q}\in\C_N\p{2}}
+If $M=1$, then
+$\p{\m h\circ\m w}_k=\sum_{\ell\in\ZZ_N} h_{k\ell} w_\ell$ and
+$\tilde h_k = h_k \equiv \sum_{\ell\in\ZZ_N} h_{k\ell}$, so
+\begin{align}
+  H_0 X_{\m w} \ket\psi
+  = E_0 X_{\m w} \ket\psi + \sum_{k\in\ZZ_N}
+  \p{\sum_{\ell\in\ZZ_N} h_{k\ell} w_\ell - h_k w_k} X_k \ket\psi,
+\end{align}
+which is precisely what we found in
+\eqref{eq:single_body_diagnosis_end}.  If $M=2$, then
+$\p{\m h\circ\m w}_{pq} = \sum_{k\in\ZZ_N}
+\p{h_{pk}w_{kq}+h_{qk}w_{pk}}$, so
+\begin{align}
+  H_0 X_{\m w} \ket\psi
+  = E_0 X_{\m w} \ket\psi + \sum_{\p{p,q}\in\C_N\p{2}}
   \sp{\sum_{k\in\ZZ_N}\p{h_{pk} w_{kq} + h_{qk} w_{pk}}
     - \tilde h_{pq} w_{pq}}
-  O_k \ket\psi,
-\end{align}
-where $\tilde h_{pq}=h_p+h_q-2h_{pq}$ with $h_p\equiv\sum_qh_{pq}$.
+  X_k \ket\psi,
+\end{align}
+where $\tilde h_{pq}=h_p+h_q-2h_{pq}$.
 
@@
 \subsection{Translational invariance, isotropy, and other symmetries}
-\label{sec:multi_body_symmetries}
-
-Solving the eigenvalue problem in \eqref{eq:H_VM_psi_eig} to find
-$M$-body perturbations $\V_M$ that generate states of definite
+\label{sec:symmetries}
+
+Solving the eigenvalue problem in \eqref{eq:multi_body_eig} to find
+$M$-body perturbations $X_{\m w}$ that generate states of definite
 interaction energy nominally requires diagonalizing the matrix
-$\check{\m h}$ with dimensions
+$\hat{\m h}$ with dimensions
 ${N\choose M}\times{N\choose M}\sim N^M\times N^M$.  If the
-interaction Hamiltonian $H_0$ and multi-body perturbation $\V_M$ have
-any symmetries, such as translational invariance isotropy, then we can
-use these symmetry to reduce the eigenvalue problem in
-\eqref{eq:H_VM_psi_eig}.  Translational invariance, for example, can
-reduce this problem to diagonalizing a matrix with dimensions
+interaction Hamiltonian $H_0$ and multi-body perturbation $X_{\m w}$
+have any symmetries, such as translational invariance isotropy, then
+we can use these symmetries to reduce the complexity of this
+eigenvalue problem.  Translational invariance, for example, can reduce
+this problem to that of diagonalizing a matrix with dimensions
 $\sim{N-1\choose M-1}\times{N-1\choose M-1}\sim N^{M-1}\times
@@
 that depend on the lattice geometry.  The complexity of diagonalizing
-a matrix with dimensions $d\times d$ is $\O\p{d^3}$, so even a
+a matrix with dimensions $d\times d$ is $O\p{d^3}$, so even a
 reduction of $d$ by a factor of 2 yields practically significant gains
@@
 provide a general prescription for exploiting the symmetries of $H_0$
-and $\V_M$ to simplify the eigenvalue problem in
-\eqref{eq:H_VM_psi_eig}.
-
-Symmetries of $H_0$ and $\V_M$ essentially break the set of choices
-$\C_N\p{M}$ into equivalence classes $\S_M$\footnote{For brevity, we
-  will suppress the explicit dependence of $\S_M$ on $N$, as well as
-  the dependence of $\S_M$ on any other factors such as lattice
-  geometry.}, where elements $s\in\S_M$ are subsets
+and $X_{\m w}$ to simplify the eigenvalue problem in
+\eqref{eq:multi_body_eig}.
+
+Symmetries of $H_0$ and $X_{\m w}$ essentially break the set of
+choices $\C_N\p{M}$ into equivalence classes $\S_M$\footnote{For
+  brevity, we will suppress the explicit dependence of $\S_M$ on $N$,
+  as well as the dependence of $\S_M$ on any other factors such as
+  lattice geometry.}, where elements $s\in\S_M$ are subsets
 $s\subset\C_N\p{M}$ that are invariant under the appropriate symmetry
@@
 Our task is essentially to express the multi-body eigenvalue problem
-in \eqref{eq:H_VM_psi_eig} in a basis that obeys the symmetries under
-consideration.  To this end, for any equivalence class $s\in\S_M$ we
-define a uniform superposition $\ket{s}$ over all choices $k\in s$, as
-well as the projector $\P_\S$ onto the space of uniform
-superpositions:
+in \eqref{eq:multi_body_eig} in a basis that obeys the symmetries
+under consideration.  To this end, for any equivalence class
+$s\in\S_M$ we define a uniform superposition $\ket{s}$ over all
+choices $k\in s$, as well as the projector $\P_\S$ onto the space of
+uniform superpositions:
 \begin{align}
@@
 where $w_s\equiv w_k$ for any $k\in s$.  Finally, we define the
-projection of $\check{\m h}$ onto the symmetric subspace:
-\begin{align}
-  \check{\m h}_{\S_M}
-  \equiv \P_{\S_M} \check{\m h} \P_{\S_M}
+projection of $\hat{\m h}$ onto the symmetric subspace:
+\begin{align}
+  \hat{\m h}_{\S_M}
+  \equiv \P_{\S_M} \hat{\m h} \P_{\S_M}
   = \sum_{s\in\S_M} \ket{s} \sp{
@@
 \begin{align}
-  \check{\m h}_{\S_M} \c \v{\m w} = \Delta \v{\m w}.
+  \hat{\m h}_{\S_M} \c \v{\m w} = \Delta \v{\m w}.
 \end{align}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Projecting products of operators onto the permutationally
-  symmetric manifold}
-\label{sec:sym_prod}
-
-Weak $M$-body perturbations $\V_M$ to SU($n$)-symmetric interactions
-can be treated perturbatively, resulting in an effective Hamiltonian
-on a permutationally symmetric manifold $\M_0$ of $N$ spins.
-Computing this effective Hamiltonian generally requires multiplying
-several operators, and projecting the resulting product onto $\M_0$.
-Even if the perturbation $\V_M$ is too strong for rigorous
-perturbative treatment, as long as it is comparable in magnitude to
-the energetic penalty incurred by breaking permutational symmetry, its
-effects can still be accounted for by an effective theory on the joint
-manifold of permutationally symmetric states together with states of
-``weak'' permutational symmetry breaking, e.g.~the manifold of states
-generated by the action of $M$-body perturbations on permutationally
-symmetric states.  The expansion of operators in a basis for this
-extended manifold likewise requires multiplying several operators and
-projecting the resulting product onto $\M_0$.  Specifically, if
-$\V_M=\sum_{k\in\C_N\p{M}}w_kO_k$, then we can define operators
-\begin{align}
-  \V_M^{(k)} \equiv \sum_{p\in\C_N\p{M}} w^{(k)}_p O_p,
-\end{align}
-where $\set{w^{(k)}}$ forms a complete basis for the set of symmetric
-tensors of dimension $M$ with zeros on all diagonals (i.e.~zero in all
-entries with a repeated index).  Denoting a complete basis of
-permutationally symmetric states by $\set{\ket{m}}$, we can then
-identify the $M$-body perturbed states
-\begin{align}
-  \ket{mk} \equiv \f1{\sqrt{\N_{mk}}} \V_M^{(k)} \ket{m},
-  &&
-  \N_{mk} \equiv \bk{m | \V_M^{(k)} \V_M^{(k)} | m}.
-\end{align}
-The restriction of a multi-body operator $\Q=\sum_p v_p Q_p$ to this
-expanded Hilbert space is then
-\begin{align}
-  \Q =_M \P_M \Q \P_M,
-  &&
-  \P_M \equiv \sum_m \op{m} + \sum_{m,k} \op{mk},
-\end{align}
-where $=_M$ denotes equality up to a restriction to the support of
-$\P_M$, and finding matrix elements of $\Q$ now requires computing
-products of the form
-\begin{align}
-  \bk{\ell | \V_M^{(j)} \Q \V_M^{(k)} | m}
-  = \sum_{p,q,r} w^{(j)}_p v_q w^{(k)}_r \bk{\ell | O_p Q_q O_r | m}.
-  \label{eq:example_prod}
-\end{align}
-The purpose of this section is to work out such products in a
-systematic way.  To this end, we derive a general diagrammatic
-expansion for these products, as well as a set of rules to simplify
-these expansions so as to minimize their computational cost.  These
-diagrammatic methods allow us to algorithmically evaluate products
-such as \eqref{eq:example_prod} from the bare multi-body operator data
-$\p{w^{(j)},O},\p{v,Q},\p{w^{(k)},O}$.  We then work out some special
-cases of products that are relevant to the perturbation theory in
-Sections \ref{sec:single_body_pert} and \ref{sec:two_body_pert}.
+\section{Projecting products of multi-body operators onto the
+  permutationally symmetric manifold}
+\label{sec:operator_product}
+
+\red{[TODO: write introductory text for this section]}
 
@@
 Given a system of $N$ spins, we wish to project a product of
-multi-body operators onto the permutationally symmetric manifold.  For
-$p$ multi-body operators, the projection of such a product takes the
-form
+multi-body operators onto the permutationally symmetric manifold
+$\M_0$.  For $p$ multi-body operators, the projection of such a
+product takes the form
 \begin{align}
   \P_0 \sp{\prod_{j\in\ZZ_p}
-    \sum_{k\in\C_N\p{M_j}} w_j\p{k} O_j\p{k}} \P_0,
+    \sum_{k\in\C_N\p{M_j}} w_j\p{k} X_j\p{k}} \P_0,
   \label{eq:sym_prod_proj}
 \end{align}
-where $\P_0$ is a projector onto the permutationally symmetric
-manifold; $k\in\C_N\p{M_j}$ is a choice of $M_j$ distinct spins;
-$O_j\p{k}$ denotes the action of an $M_j$-spin operator on spins $k$
-that is invariant under permutations of $k$; and $w_j\p{k}$ is a
-scalar coefficient for $O_j\p{k}$.
+where $\P_0$ is a projector onto the PS manifold $\M_0$; $j$ indexes
+one of the $p$ multi-body operators; $X_j$ is an $M_j$-spin operator
+that is invariant under arbitrary permutation of its tensor factors;
+$k\in\C_N\p{M_j}$ is a choice of $M_j$ spins that $X_j\p{k}$ acts on;
+$\C_N\p{M_j}$ is the set of all choices of $M_j$ elements from a set
+of size $N$; and $\m w_j$ are dimension-$M_j$ tensors with scalar
+entries $w_j\p{k}$. \red{[TODO: relax the restrictions on $X_j$ and
+  $\m w_j$ to allow for more general perturbations, e.g.~of the form
+  $s_\x\otimes s_\y-s_\y\otimes s_\x$]}
 
@@
 \begin{align}
-  \prod_{j\in\ZZ_p} \sum_{k\in\C_N\p{M_j}} w_j\p{k} O_j\p{k}
+  \prod_{j\in\ZZ_p} \sum_{k\in\C_N\p{M_j}} w_j\p{k} X_j\p{k}
   = \prod_{j\in\ZZ_p} \f1{M_j!} \sum_{k\in\ZZ_N^{M_j}}
-  w_j\p{k} O_j\p{k},
+  w_j\p{k} X_j\p{k},
   \label{eq:sym_prod_vec}
 \end{align}
-where we extend the definitions of $w_j\p{k}$ and $O_j\p{k}$ to all
-vectors $k\in\ZZ_N^{M_j}$ by defining symmetric dimension-$M_j$
-tensors $w_j$ and $O_j$ (i.e.~tensors with $M_j$ indices, with $O_j$
-an {\it operator-valued} tensor) for which $w_j\p{k}=O_j\p{k}=0$ if
-$k$ contains any repeated symbols.  We then collect terms whose
-operator content is equivalent up to a permutation of spins.  To this
-end, we classify terms in \eqref{eq:sym_prod_vec} by the numbers $g_S$
-of indices shared by all tensors $w_j,O_j$ with $j\in S\subset\ZZ_p$.
-For example, a term of the form
+where we extend the definitions of $w_j\p{k}$ and $X_j\p{k}$ to all
+vectors $k\in\ZZ_N^{M_j}$ by enforcing that these objects are
+symmetric with respect to permutations of the entries in $k$, and that
+$w_j\p{k}=X_j\p{k}=0$ if $k$ contains any repeated entries.  We then
+collect terms whose operator content is equivalent up to a permutation
+of spins.  To this end, we classify terms in \eqref{eq:sym_prod_vec}
+by the numbers $g_S$ of indices shared by all tensors $w_j$ with
+$j\in S\subset\ZZ_p$.  For example, a term of the form
 $w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}$ with distinct indices
@@
 \begin{align}
-  \prod_{j\in\ZZ_p} \sum_{k\in\ZZ_N^{M_j}} w_j\p{k} O_j\p{k}
-  \EQPS \sum_{g\in\G\p{\m M}} \S\p{g} w\p{g} O\p{g},
+  \prod_{j\in\ZZ_p} \sum_{k\in\ZZ_N^{M_j}} w_j\p{k} X_j\p{k}
+  \EQPS \sum_{g\in\G\p{\m M}} \S\p{g} w\p{g} X\p{g},
   \label{eq:sym_prod_group_start}
@@
 where $\EQPS$ denotes equality up to a restriction to the
-permutationally symmetric manifold; $O\p{g}$ is an operator acquired
-by assigning indices to the tensors $O_j$ in a manner consistent with
-$g$; $w\p{g}$ is a scalar acquired by summing over all indices
-assigned to the tensors $w_j$ according to $g$; and $\S\p{g}$ is a
-symmetry factor accounting for the number of ways to assign indices
-according to $g$.
+permutationally symmetric manifold; $X\p{g}$ is an operator acquired
+by assigning indices (choices of spins) to all $X_j$ in a manner
+consistent with $g$; $w\p{g}$ is a scalar acquired by summing over all
+indices assigned to the tensors $w_j$ according to $g$; and $\S\p{g}$
+is a symmetry factor accounting for the number of ways to assign
+indices according to $g$.
 
@@
 index values, we define the restriction of $k\in\C_N\p{g}$ to values
-associated with $w_j$ and $O_j$:
+associated with $w_j$ and $X_j$:
 \begin{align}
@@
   &&
-  O\p{g} \equiv \prod_{j\in\ZZ_p}
-  O_j\p{\ell_j}~\t{for some}~\ell\in\C_N\p{g},
+  X\p{g} \equiv \prod_{j\in\ZZ_p}
+  X_j\p{\ell_j}~\t{for some}~\ell\in\C_N\p{g},
   \label{eq:diagram_factors}
@@
 where the choice of $\ell\in\C_N\p{g}$ does not matter after
-projection onto permutationally symmetric manifold, and
+projection onto permutationally symmetric manifold.  Finally, we can
+work out that
 \begin{align}
@@
 that account for the number of ways to partition the indices of each
-$w_j,O_j$ into sets of shared indices as specified by $g$, and the
-second product in \eqref{eq:sym_prod_symmetry} accounts for the number
-of ways to permute each set of shared indices on every tensor.
+$w_j$ into sets of shared indices as specified by $g$, and the second
+product in \eqref{eq:sym_prod_symmetry} accounts for the number of
+ways to permute each set of shared indices on every tensor.
 Altogether, the symmetry factors $\S\p{g}$ cancel out with the
@@
 \begin{align}
-  \prod_{j\in\ZZ_p} \sum_{k\in\C_N\p{M_j}} w_j\p{k} O_j\p{k}
-  \EQPS \sum_{g\in\G\p{\m M}} w\p{g} O\p{g}.
+  \prod_{j\in\ZZ_p} \sum_{k\in\C_N\p{M_j}} w_j\p{k} X_j\p{k}
+  \EQPS \sum_{g\in\G\p{\m M}} w\p{g} X\p{g}.
   \label{eq:sym_prod_group}
 \end{align}
+This result essentially gives us a diagrammatic expansion for the
+product of multi-body operators, projected onto the permutationally
+symmetric manifold.
 
@@
 $a,b,c,d,e$ are distinct.  Such a diagram with $\abs{g}$ dots
-nominally takes $\O\p{N^{\abs{g}}}$ time to compute.  In practice,
-this computational cost can be prohibitive for performing numerical
+nominally takes $O\p{N^{\abs{g}}}$ time to compute.  In practice, this
+computational cost can be prohibitive for performing numerical
 simulations of systems with any appreciable size.  We therefore wish
@@
 $a\in\ZZ_N$ and $a\in\set{b,c}$.  The sum over $a\in\ZZ_N$ to compute
-$v\p{\circ,b}$ has $\O\p{N^2}$ cost, and can be performed prior to
+$v\p{\circ,b}$ has $O\p{N^2}$ cost, and can be performed prior to
 evaluating the diagram in \eqref{eq:example_elim_o}.  The
 decomposition in \eqref{eq:example_elim} therefore splits one
-$\O\p{N^3}$ diagram into two $\O\p{N^2}$ diagrams.
+$O\p{N^3}$ diagram into two $O\p{N^2}$ diagrams.
 
@@
 \subsubsection{Symmetry factors}
+\label{sec:symmertry_factors}
 
@@
 automate the evaluation of \eqref{eq:sym_prod_group} is expansion of
-the multi-spin operators $O\p{g}$ in a basis for the permutationally
+the multi-spin operators $X\p{g}$ in a basis for the permutationally
 symmetric manifold $\M_0$.  Without loss of generality, we consider a
-single permutationally symmetric $M$-spin operator $Q$; our task is
+single permutationally symmetric $M$-spin operator $X$; our task is
 essentially to find coefficients of the expansion
 \begin{align}
-  \P_0 Q \P_0
-  = \sum_{a,b} \bk{a|Q|b} \op{a}{b},
+  \P_0 X \P_0
+  = \sum_{a,b} \bk{a|X|b} \op{a}{b},
 \end{align}
 where $\set{\ket{a}},\set{\ket{b}}$ are complete bases for $\M_0$, and
-the choice of $M$ spins on which $Q$ acts is arbitrary.  In a system
+the choice of $M$ spins on which $X$ acts is arbitrary.  In a system
 of $N$ spins with $n$ states each, a suitable basis for $\M_0$ can be
@@
 \begin{align}
-  \bk{a|Q|b}
+  \bk{a|X|b}
   = \sum_{\substack{\alpha,\beta\in\A_n\p{M}\\\alpha\le a,\,\beta\le b}}
@@
     {\C\p{a}\C\p{b}}}
-  \bk{\alpha|Q|\beta},
+  \bk{\alpha|X|\beta},
   \label{eq:multi_body_eval}
@@
 $\beta=b-a+\alpha$.  Note that, by slight abuse of notation, the
-operator $Q$ on the left of \eqref{eq:multi_body_eval} acts on an
-arbitrary choice of $M$ spins (out of $N$), whereas the operator $Q$
+operator $X$ on the left of \eqref{eq:multi_body_eval} acts on an
+arbitrary choice of $M$ spins (out of $N$), whereas the operator $X$
 on the right of \eqref{eq:multi_body_eval} is simply an $M$-spin
-operator, with matrix elements $\bk{\alpha|Q|\beta}$ evaluated with
+operator, with matrix elements $\bk{\alpha|X|\beta}$ evaluated with
 respect to the $M$-spin states $\ket\alpha,\ket\beta$.
@@
 \subsection{Two single-body operators}
-\label{sec:PXYP}
-
-Here we simplify the product
-\begin{align}
-  \P_0 \sp{\sum_{p,q} v_p w_q X_p Y_q} \P_0
+\label{sec:single_pair_prod}
+
+Here we simplify the single-body product
+\begin{align}
+  \P_0 X_{\m v} Y_{\m w} \P_0
+  = \P_0 \sp{\sum_{p,q\in\ZZ_N} v_p w_q X_p Y_q} \P_0
   \EQPS \diagram{single_body_0} X_1 Y_2
   + \diagram{single_body_1} X_1 Y_1,
-  \label{eq:PXYP_start}
 \end{align}
@@
 $H_1^{(2)}$ induced on the permutationally symmetric manifold $\M_0$
-by the single-body perturbation $\V_1$ in \eqref{eq:perturbations}.
-Defining
-\begin{align}
-  \v m \equiv \sum_p m_p \ket{p},
-  &&
-  \col{m} \equiv \sum_p m_p,
-\end{align}
-for $m\in\set{v,w}$, we can simplify
-\begin{align}
-  \diagram{single_body_1}
-  \equiv \sum_p v_p w_p
-  = \v v\c\v w,
+by a single-body perturbation.  Defining
+\begin{align}
+  \v{\m m} \equiv \sum_{p\in\ZZ_N} m_p \ket{p},
+  &&
+  \col{m} \equiv \sum_{p\in\ZZ_N} m_p,
+\end{align}
+for both $\m m\in\set{\m v,\m w}$, we can simplify
+\begin{align}
+  \diagram{single_body_1} = \v{\m v}\c\v{\m w},
 \end{align}
@@
   = \diagram{single_body_0_oo} - \diagram{single_body_1}
-  = \col{u}\,\col{v} - \v v \c\v w,
+  = \col{v}\,\col{w} - \v{\m v} \c\v{\m w},
 \end{align}
@@
 \begin{align}
-  \sum_{p,q} v_p w_q X_p Y_q
-  \EQPS \col{u}\,\col{v} X_1 Y_2 - \v v\c\v w \p{X_1 Y_2 - X_1 Y_1},
+  \sum_{p,q\in\ZZ_N} v_p w_q X_p Y_q
+  \EQPS \col{v}\,\col{w} X_1 Y_2
+  - \v{\m v}\c\v{\m w} \p{X_1 Y_2 - X_1 Y_1},
 \end{align}
 In order to write this result in terms of collective operators
-$\col{Z} \equiv \sum_p Z_p$, we expand
+$\col{Z} \equiv \sum_{p\in\ZZ_N} Z_p$, we expand
 \begin{align}
@@
 \begin{align}
-  \sum_{p,q} v_p w_q X_p Y_q
-  \EQPS \f{\col{u}\,\col{v}}{N\p{N-1}}
+  \sum_{p,q\in\ZZ_N} v_p w_q X_p Y_q
+  \EQPS \f{\col{v}\,\col{w}}{N\p{N-1}}
   \p{\col{X}\,\col{Y} - \col{XY}}
-  - \f{\v v \c\v w}{N\p{N-1}}
+  - \f{\v{\m v}\c\v{\m w}}{N\p{N-1}}
   \p{\col{X}\,\col{Y} - N\col{XY}}.
@@
 \subsection{Two two-body operators}
-\label{sec:POQP}
-
-Here we simplify the product
+\label{sec:two_pair_prod}
+
+Here we simplify the two-body product
 \begin{multline}
-  \P_0 \sp{\sum_{k<\ell} \sum_{p<q}
-    v_{k\ell} w_{pq} O_{k\ell} Q_{pq}} \P_0 \\
-  \EQPS \diagram{two_body_0} O_{1,2} Q_{3,4}
-  + \diagram{two_body_1} O_{1,2} Q_{1,3}
-  + \diagram{two_body_2} O_{1,2} Q_{1,2},
-  \label{eq:POQP_start}
+  \P_0 X_{\m v} Y_{\m w} \P_0
+  = \P_0 \sp{\sum_{\p{k,\ell},\p{p,q}\in\C_N\p{2}}
+    v_{k\ell} w_{pq} X_{k\ell} Y_{pq}} \P_0 \\
+  \EQPS \diagram{two_body_0} X_{1,2} Y_{3,4}
+  + \diagram{two_body_1} X_{1,2} Y_{1,3}
+  + \diagram{two_body_2} X_{1,2} Y_{1,2},
 \end{multline}
@@
 $v_{k\ell},w_{pq}$ are scalar coefficients with $m_{pq}=m_{qp}$ and
-$m_{pp}=0$ for each of $m\in\set{v,w}$; and $O,Q$ are permutationally
-symmetric two-spin operators.  This product appears, for example, in
-the calculation of the second-order effective Hamiltonian $H_2^{(2)}$
-induced on the permutationally symmetric manifold $\M_0$ by the
-two-body perturbation $\V_2$ in \eqref{eq:perturbations} in
-\eqref{eq:perturbations}.  Defining
-\begin{align}
-  \v{\m m} \equiv \sum_{p<q} m_{pq} \ket{\set{p,q}},
-  &&
-  \v m \equiv \sum_p m_p \ket{p},
-  &&
-  m_p \equiv \sum_q m_{pq},
-  &&
-  \col{m} \equiv \sum_{p<q} m_{pq},
-\end{align}
-for $m\in\set{v,w}$, we can write
-\begin{align}
-  \diagram{two_body_2} \equiv \sum_{\p{p,q}\in\C_N\p{2}} v_{pq} w_{pq}
-  = \v{\m v} \c\v{\m w},
+$m_{pp}=0$ for both $\m m\in\set{\m v,\m w}$; and $X,Y$ are
+permutationally symmetric two-spin operators.  This product appears,
+for example, in the calculation of the second-order effective
+Hamiltonian $H_2^{(2)}$ induced on the permutationally symmetric
+manifold $\M_0$ by a two-body perturbation.  Defining
+\begin{align}
+  \v{\m m} \equiv \sum_{\p{p,q}\in\C_N\p{2}} m_{pq} \ket{\set{p,q}},
+  &&
+  \v m \equiv \sum_{p,q\in\ZZ_N} m_{pq} \ket{p},
+  &&
+  \col{m} \equiv \sum_{\p{p,q}\in\C_N\p{2}} m_{pq},
+\end{align}
+for both $\m m\in\set{\m v,\m w}$, we can write
+\begin{align}
+  \diagram{two_body_2} = \v{\m v} \c\v{\m w},
 \end{align}
@@
 Altogether,
+\begin{align}
+  X_{\m v} Y_{\m w}
+  \EQPS \p{\col{v}\,\col{w} -\v v\c\v w + \v{\m v}\c\v{\m w}}
+  X_{1,2} Y_{3,4}
+  + \p{\v v\c\v w - 2 \v{\m v} \c \v{\m w}} X_{1,2} Y_{1,3}
+  + \v{\m v} \c \v{\m w}\, X_{1,2} Y_{1,2}.
+\end{align}
+In the special case that $X=Y=Z\otimes Z$ with a single-spin operator
+$Z$ for which $Z^2=1$, we can further simplify
+\begin{align}
+  \sum_{\p{k,\ell},\p{p,q}\in\C_N\p{2}} v_{k\ell} w_{pq}
+  Z_k Z_\ell Z_p Z_q
+  \EQPS \p{\col{v}\,\col{w} -\v v\c\v w + \v{\m v}\c\v{\m w}}
+  Z^{\otimes 4}
+  + \p{\v v\c\v w - 2 \v{\m v} \c \v{\m w}} Z^{\otimes 2}
+  + \v{\m v}\c\v{\m w}.
+\end{align}
+In terms of collective operators
+$\col{Z} \equiv \sum_{p\in\ZZ_N} Z_p$, we then have
 \begin{multline}
-  \sum_{k<\ell} \sum_{p<q} v_{k\ell} w_{pq} O_{k\ell} Q_{pq} \\
-  \EQPS \p{\col{v}\,\col{w} -\v v\c\v w + \v{\m v}\c\v{\m w}}
-  O_{1,2} Q_{3,4}
-  + \p{\v v\c\v w - 2 \v{\m v} \c \v{\m w}} O_{1,2} Q_{1,3}
-  + \v{\m v} \c \v{\m w}\, O_{1,2} Q_{1,2}.
-  \label{eq:POQP}
-\end{multline}
-In the special case that $O=Q=X\otimes X$ with $X^2=1$, we can further
-simplify
-\begin{align}
-  \sum_{k<\ell} \sum_{p<q} v_{k\ell} w_{pq} X_k X_\ell X_p X_q
-  \EQPS
-  \p{\col{v}\,\col{w} -\v v\c\v w + \v{\m v}\c\v{\m w}} X^{\otimes 4}
-  + \p{\v v\c\v w - 2 \v{\m v} \c \v{\m w}} X^{\otimes 2}
-  + \v{\m v}\c\v{\m w}.
-\end{align}
-In terms of collective operators $\col{X} \equiv \sum_p X_p$, we then
-have
-\begin{multline}
-  \sum_{k<\ell} \sum_{p<q} v_{k\ell} w_{pq} X_k X_\ell X_p X_q \\
-  \EQPS
-  \f{\col{v}\,\col{w} -\v v\c\v w + \v{\m v}\c\v{\m w}}{N!_4}
-  \sp{\col{X}^4 - 2\p{3N-4} \col{X}^2}
-  + \f{\p{\v v\c\v w - 2 \v{\m v} \c \v{\m w}}}{N!_2}\, \col{X}^2 \\
-  + \f{3\,\col{v}\,\col{w} - N \v v\c\v w
-    + N\p{N-2}\v{\m v}\c\v{\m w}}{\p{N-1}\p{N-3}},
+  \sum_{\p{k,\ell},\p{p,q}\in\C_N\p{2}} v_{k\ell} w_{pq}
+  Z_k Z_\ell Z_p Z_q \\
+  \EQPS \f{\col{v}\,\col{w} -\v v\c\v w + \v{\m v}\c\v{\m w}}{N!_4}
+  \sp{\col{Z}^4 - 2\p{3N-4} \col{Z}^2}
+  + \f{\p{\v v\c\v w - 2 \v{\m v} \c \v{\m w}}}{N!_2}\, \col{Z}^2 \\
+  + \f{3\,\col{v}\,\col{w} - N \v v\c\v w + N\p{N-2}\v{\m v}\c\v{\m
+      w}}{\p{N-1}\p{N-3}},
 \end{multline}
@@
 \begin{align}
-  \X \equiv \P_0 \sp{\sum_{k<\ell} \sum_{p<q} \sum_{r<s}
-    u_{k\ell} v_{pq} w_{rs} X_k X_\ell X_p X_q X_r X_s} \P_0,
-  &&
-  X^2 = 1,
+  \Z \equiv \P_0 \sp{\sum_{\p{k,\ell},\p{p,q},\p{r,s}\in\C_N\p{2}}
+    u_{k\ell} v_{pq} w_{rs} Z_k Z_\ell Z_p Z_q Z_r Z_s} \P_0,
+  &&
+  Z^2 = 1,
 \end{align}
 where $\P_0$ is a projector onto the permutationally symmetric
-manifold $\M_0$; $k,\ell,p,q,r,s$ index individual spins; $X$ is a
+manifold $\M_0$; $k,\ell,p,q,r,s$ index individual spins; $Z$ is a
 single-spin operator; and $u_{k\ell},v_{pq},w_{rs}$ are scalar
 coefficients with $m_{pq}=m_{qp}$ and $m_{pp}=0$ for each of
-$m\in\set{u,v,w}$.  Collecting terms according to the number of $X$
-operators that remain after accounting for $X^2=1$, we find that
-\begin{align}
-  \X \EQPS A_6 X^{\otimes 6} + A_4 X^{\otimes 4}
-  + A_2 X^{\otimes 2} + A_0,
+$\m m\in\set{\m u,\m v,\m w}$.  \red{[TODO: give example of when this
+  product appears.]} Collecting terms according to the number of $Z$
+operators that remain after accounting for $Z^2=1$, we find that
+\begin{align}
+  \Z \EQPS A_6 Z^{\otimes 6} + A_4 Z^{\otimes 4}
+  + A_2 Z^{\otimes 2} + A_0,
   \label{eq:triple_multi}
@@
 \end{align}
-The first diagram in each of $A_6,A_4,A_2,A_0$ as written in
+The last diagram in each of $A_6,A_4,A_2,A_0$ as written in
 \eqref{eq:triple_64}, \eqref{eq:triple_2} thus has only one assignment
@@
 $A_6,A_4,A_2,A_0$, we now need to simplify all diagrams to reduce
-their computational complexity, as discussed in Section
-\ref{sec:diagrams}.  Simplifying all these diagrams by hand is
-tedious, but the process of eliminating filled dots and crosses from a
-diagram is algorithmic enough to execute on a computer, doing which we
-find
+their computational complexity, as discussed in Appendix
+\ref{sec:diagrams}.  The process of eliminating filled dots and
+crosses from a diagram is algorithmic enough to execute on a computer,
+doing which we find
 \begin{align}
@@
 \end{align}
-Defining, for each of $m\in\set{u,v,w}$,
-\begin{align}
-  \m m \equiv \sum_{p,q} m_{pq} \op{p}{q},
-  &&
-  \v{\m m} \equiv \sum_{p<q} m_{pq} \ket{\set{p,q}},
-  &&
-  \v m \equiv \sum_p m_p \ket{p},
-  &&
-  m_p \equiv \sum_q m_{pq},
-  &&
-  \col{m} \equiv \sum_{p<q} m_{pq},
+Defining, for each of $\m m\in\set{\m u,\m v,\m w}$,
+\begin{align}
+  \v{\m m} \equiv \sum_{\p{p,q}\in\C_N\p{2}} m_{pq} \ket{\set{p,q}},
+  &&
+  \v m \equiv \sum_{p\in\ZZ_N} m_p \ket{p},
+  &&
+  m_p \equiv \sum_{q\in\ZZ_N} m_{pq},
+  &&
+  \col{m} \equiv \sum_{\p{p,q}\in\C_N\p{2}} m_{pq},
 \end{align}
@@
   D_4 = \col{u}\,\v v\c\v w + \col{v}\,\v w\c\v u
-  + \col{w}\,\v u\c\v v - 2 \sum_p u_p v_p w_p,
+  + \col{w}\,\v u\c\v v - 2 \sum_{p\in\ZZ_N} u_p v_p w_p,
   &&
@@
   + \col{w}\,\v{\m u}\c\v{\m v} \\
-  - 2 \sum_{p,q} \p{u_p v_{pq} w_{pq}
+  - 2 \sum_{p,q\in\ZZ_N} \p{u_p v_{pq} w_{pq}
     + v_p w_{pq} u_{pq} + w_p u_{pq} v_{pq}}
-  + 4 \sum_{p<q} u_{pq} v_{pq} w_{pq}.
+  + 4 \sum_{\p{p,q}\in\C_N\p{2}} u_{pq} v_{pq} w_{pq}.
 \end{multline}
-Finally, we can also write the product $\X$ in \eqref{eq:triple_multi}
-in terms of the collective operator $\col{X} \equiv \sum_p X_p$:
-\begin{align}
-  \X \EQPS
-  \tilde A_6 \col{X}^6 + \tilde A_4 \col{X}^4
-  + \tilde A_2 \col{X}^2 + \tilde A_0,
+Finally, we can also write the product $\Z$ in \eqref{eq:triple_multi}
+in terms of the collective operator
+$\col{Z} \equiv \sum_{p\in\ZZ_N} Z_p$:
+\begin{align}
+  \Z \EQPS
+  \tilde A_6 \col{Z}^6 + \tilde A_4 \col{Z}^4
+  + \tilde A_2 \col{Z}^2 + \tilde A_0,
   \label{eq:triple_col}
@@
 
-\bibliography{multilevel_spin_notes.bib}
+\bibliography{\jobname.bib}
 

### 2020-03-19 collective_sun/sun_perturbation.tex
emphasized that *spatial* symmetries simplify the multi-body eigenvalue problem broke off the subsubsection about evaluating multi-spin operators into its own subsection

@@
 \newcommand{\A}{\mathcal{A}}
+\newcommand{\B}{\mathcal{B}}
 \newcommand{\C}{\mathcal{C}}
@@
 matrix of dimensions
-${N\choose M}\times{N\choose M}\sim N^M\times N^M$, but symmetries of
-$\hat{\m h}$ such as translational invariance or isotropy can
+${N\choose M}\times{N\choose M}\sim N^M\times N^M$, but spatial
+symmetries of $H_0$ such as translational invariance or isotropy can
 drastically reduce the complexity of this eigenvalue problem.  We
@@
 by examining the maximum populations within all $\M_r$ for
-$r\le r_{\t{max}}$ throughout simulation.  As long as these
-populations fall off with increasing $r$ and the maximal population
-within large-$r$ manifolds $\M_r$ is small, one should expect such
-simulations to faithfully capture (at least qualitatively) the
-dynamical behavior of the spin system under consideration.
+$r\le r_{\t{max}}$ throughout simulation.  As long as
+\begin{enumerate*}
+\item these
+populations fall off with increasing $r$,
+\item and the maximal population within large-$r$ manifolds $\M_r$ is
+  small,
+\end{enumerate*}
+one should expect such simulations to faithfully capture (at least
+qualitatively) the dynamical behavior of the spin system under
+consideration.
+
+In order to project operators onto $\ul{\M}_{r_{\t{max}}}$, we need to
+construct a basis for each $\M_r$ with $r\le r_{\t{max}}$.
+\red{[TODO: finish section]}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Translational invariance, isotropy, and other symmetries}
+\subsection{Exploiting spatial symmetries}
 \label{sec:symmetries}
@@
 ${N\choose M}\times{N\choose M}\sim N^M\times N^M$.  If the
-interaction Hamiltonian $H_0$ and multi-body perturbation $X_{\m w}$
-have any symmetries, such as translational invariance isotropy, then
-we can use these symmetries to reduce the complexity of this
+interaction Hamiltonian $H_0$ has any spatial symmetries, however,
+then we can use these symmetries to reduce the complexity of this
 eigenvalue problem.  Translational invariance, for example, can reduce
 this problem to that of diagonalizing a matrix with dimensions
-$\sim{N-1\choose M-1}\times{N-1\choose M-1}\sim N^{M-1}\times
-N^{M-1}$, and isotropy can further reduce the problem size by factors
-that depend on the lattice geometry.  The complexity of diagonalizing
-a matrix with dimensions $d\times d$ is $O\p{d^3}$, so even a
-reduction of $d$ by a factor of 2 yields practically significant gains
-(a factor of $2^3=8$) in the runtime of diagonalization.  Here, we
-provide a general prescription for exploiting the symmetries of $H_0$
-and $X_{\m w}$ to simplify the eigenvalue problem in
-\eqref{eq:multi_body_eig}.
-
-Symmetries of $H_0$ and $X_{\m w}$ essentially break the set of
+$\le{N-1\choose M-1}\times{N-1\choose M-1}\sim N^{M-1}\times N^{M-1}$,
+and isotropy can further reduce the problem size by factors that
+depend on the lattice geometry.  The complexity of diagonalizing a
+matrix with dimensions $d\times d$ is $O\p{d^3}$, so even a reduction
+of $d$ by a factor of 2 yields practically significant gains (a factor
+of $2^3=8$) in the runtime of diagonalization.  Here, we provide a
+general prescription for exploiting the spatial symmetries of $H_0$ to
+simplify the eigenvalue problem in \eqref{eq:multi_body_eig}.
+
+Our task is essentially to express the multi-body eigenvalue problem
+in \eqref{eq:multi_body_eig} in a basis that respects the spatial
+symmetries of $H_0$.  These symmetries essentially break the set of
 choices $\C_N\p{M}$ into equivalence classes $\S_M$\footnote{For
@@
 
-Our task is essentially to express the multi-body eigenvalue problem
-in \eqref{eq:multi_body_eig} in a basis that obeys the symmetries
-under consideration.  To this end, for any equivalence class
-$s\in\S_M$ we define a uniform superposition $\ket{s}$ over all
+For simplicity, we first assume that the $M$-body operator $X_{\m w}$
+obeys the same spatial symmetries as $H_0$.  For any equivalence class
+$s\in\S_M$, we then define a uniform superposition $\ket{s}$ over all
 choices $k\in s$, as well as the projector $\P_\S$ onto the space of
@@
 \end{align}
+More generally, even if the couplings $\v{\m w}$ of $X_{\m w}$ do not
+obey the same spatial symmetries as $H_0$, these symmetries generally
+endow $\check{\m h}$ with a block-diagonal structure. \red{[TODO:
+  finish discussion/derivation]}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Matrix elements of individual multi-spin operators}
+\subsection{Matrix elements of individual multi-spin operators}
 
@@
   \P_0 X \P_0
-  = \sum_{a,b} \bk{a|X|b} \op{a}{b},
-\end{align}
-where $\set{\ket{a}},\set{\ket{b}}$ are complete bases for $\M_0$, and
-the choice of $M$ spins on which $X$ acts is arbitrary.  In a system
-of $N$ spins with $n$ states each, a suitable basis for $\M_0$ can be
+  = \sum_{a,b\in\B_0} \bk{a|X|b} \op{a}{b},
+\end{align}
+where $\set{\ket{a}:a\in\B_0}$ is complete bases for $\M_0$, and the
+choice of $M$ spins on which $X$ acts is arbitrary.  In a system of
+$N$ spins with $n$ states each, a suitable basis for $\M_0$ can be
 labeled by the occupation number of each single-spin state, namely
-$a=\p{a_1,a_2,\cdots,a_n}$ with each $a_j\ge 0$ and $\sum_ja_j=N$.
-Written out explicitly,
+$a=\p{a_1,a_2,\cdots,a_n}$ with each $a_\mu\ge 0$ and
+$\sum_\mu a_\mu=N$.  Written out explicitly,
 \begin{align}
@@
   &&
-  \ket{\tilde a} \equiv \bigotimes_{j=1}^n \ket{j}^{\otimes a_j},
-  &&
-  \C\p{a} \equiv \f{\p{\sum_j a_j}!}{\prod_ja_j!},
+  \ket{\tilde a} \equiv \bigotimes_\mu \ket{\mu}^{\otimes a_\mu},
+  &&
+  \C\p{a} \equiv \f{\p{\sum_\mu a_\mu}!}{\prod_\mu a_\mu!},
 \end{align}
@@
 enforces $\bk{a|a}=1$.  We denote the set of all valid assignments of
-$S$ spins to $n$ states by $\A_n\p{S}$, such that the standard basis
-for the permutationally symmetric manifold of $S$ spins is
-$\set{\ket{a}:a\in\A_n\p{S}}$.  Identifying this standard basis allows
-us to expand
+$N$ spins to $n$ states by $\A_n\p{N}$, such that the basis
+$\set{\ket{a}:a\in\A_n\p{N}}$ spans permutationally symmetric manifold
+of $N$ spins.  Identifying this basis allows us to expand
 \begin{align}
@@
 where the restriction $\alpha\le a$ and difference $a-\alpha$ are
-evaluated element-wise, i.e.~$\alpha<a\implies \alpha_j\le a_j$ and
-$\p{a-\alpha}_j=a_j-\alpha_j$ for all $j$; and $\delta_{cd}=1$ if
-$c=d$ and zero otherwise.  We sum over both $\alpha$ and $\beta$ above
-merely to keep the expression symmetric with respect to transposition;
-in practice, one can simply sum over $\alpha\in\A_n\p{M}$ and set
-$\beta=b-a+\alpha$.  Note that, by slight abuse of notation, the
-operator $X$ on the left of \eqref{eq:multi_body_eval} acts on an
-arbitrary choice of $M$ spins (out of $N$), whereas the operator $X$
-on the right of \eqref{eq:multi_body_eval} is simply an $M$-spin
-operator, with matrix elements $\bk{\alpha|X|\beta}$ evaluated with
-respect to the $M$-spin states $\ket\alpha,\ket\beta$.
+evaluated element-wise, i.e.~$\alpha<a\implies \alpha_\mu\le a_\mu$
+and $\p{a-\alpha}_\mu=a_\mu-\alpha_\mu$ for all $\mu$; and
+$\delta_{cd}=1$ if $c=d$ and zero otherwise.  We sum over both
+$\alpha$ and $\beta$ above merely to keep the expression symmetric
+with respect to transposition; in practice, one can simply sum over
+$\alpha\in\A_n\p{M}$ and set $\beta=b-a+\alpha$.  Note that, by slight
+abuse of notation, the operator $X$ on the left of
+\eqref{eq:multi_body_eval} acts on an arbitrary choice of $M$ spins
+(out of $N$), whereas the operator $X$ on the right of
+\eqref{eq:multi_body_eval} is simply an $M$-spin operator, with matrix
+elements $\bk{\alpha|X|\beta}$ evaluated with respect to the $M$-spin
+states $\ket\alpha,\ket\beta$.
 

### 2020-03-19 collective_sun/multilevel_spin_notes.tex
updating reference, giving sun_perturbation notes their own bib file

@@
 is a collective SU(2) spin-$\alpha$ operator for $\p{n-1}$ 2-level
-spins\cite{perlin2019shorttime}.  More concretely, we can embed the
+spins\cite{perlin2020shorttime}.  More concretely, we can embed the
 Hilbert space of a single $n$-level spin into the

### 2020-03-19 collective_sun/sun_perturbation.tex
started explaining the construction of operators in the 'shell basis'

@@
   \equiv \sum_{a\in\ZZ_M} \m h \circ_a \m w - \tilde{\m h} * \m w.
+  \label{eq:multi_body_eig_op}
 \end{align}
@@
 eigenvalues $\Delta$ and eigenvectors $\m m_\Delta$ for which
-$\check{\m h}\p{\m m_\Delta}=\Delta\m m_\Delta$.  Finding eigenvalues
+$\hat{\m h}\p{\m m_\Delta}=\Delta\m m_\Delta$.  Finding eigenvalues
 and eigenvectors of $\hat{\m h}$ nominally requires diagonalizing a
@@
 Defining $\m w_\Delta$ as the projection of the tensor $\m w$ onto the
-$\Delta$-eigenspace of $\check{\m h}$, we can expand
+$\Delta$-eigenspace of $\hat{\m h}$, we can expand
 \begin{align}
@@
 
-In order to project operators onto $\ul{\M}_{r_{\t{max}}}$, we need to
-construct a basis for each $\M_r$ with $r\le r_{\t{max}}$.
+In order to project an operator $X_{\m w}$ onto
+$\ul{\M}_{r_{\t{max}}}$, we first need to construct a basis for
+$\ul{\M}_{r_{\t{max}}}$, and then compute all matrix elements of
+$X_{\m w}$ in the constructed basis.  Our strategy will be to first
+identify a basis for $\M_0$, and then find, for each
+$r=1,\cdots,r_{\t{max}}$, a suitable set of $r$-body operators
+$\set{O}$ and dimension-$r$ tensors $\set{\m u}$ for which the states
+$\ket{O,\m u,m}\propto O_{\m u}\ket{m}$ form an orthonormal basis for
+$\M_r$.  Matrix elements of $X_{\m w}$ with respect to this basis for
+$\ul{\M}_{r_{\t{max}}}$ then take the form
+\begin{align}
+  \bk{O,\m u,\ell|X_{\m w}|Q,\m v,m}
+  = \f{\bk{\ell|O_{\m u}^\dag X_{\m w} Q_{\m v}|m}}
+  {\sqrt{\bk{\ell|O_{\m u}^\dag O_{\m u}|\ell}
+      \bk{m|Q_{\m v}^\dag Q_{\m v}|m}}}.
+\end{align}
+In order to compute these matrix elements, we simplify products of the
+form $\P_0 O_{\m u}^\dag X_{\m w} Q_{\m v} \P_0$ and
+$\P_0 O_{\m u}^\dag O_{\m u} \P_0$, with $\P_0$ a projector onto
+$\M_0$, in Appendix \ref{sec:operator_product}.
+
+A suitable basis for $\M_0$ is the set of PS states
+$\ket{m}=\ket{m_1,m_2,\cdots,m_n}$ with a definite occupation number
+$m_\mu$ of the single-spin state $\mu$.
+
 \red{[TODO: finish section]}
@@
   \hat{\m h}\p{\m w}
-  \equiv \p{\sum_a \m h \circ_a \m w - \tilde{\m h} * \m w}.
+  \equiv \sum_{a\in\ZZ_M} \m h \circ_a \m w - \tilde{\m h} * \m w.
   \label{eq:diagnosis_end}
@@
   \hat{\m h} \c \v{\m w} = \Delta \v{\m w},
-  \label{eq:multi_body_eig}
+  \label{eq:multi_body_eig_mat}
 \end{align}
@@
 
-Solving the eigenvalue problem in \eqref{eq:multi_body_eig} to find
-$M$-body perturbations $X_{\m w}$ that generate states of definite
-interaction energy nominally requires diagonalizing the matrix
-$\hat{\m h}$ with dimensions
+Solving the eigenvalue problem in \eqref{eq:multi_body_eig_mat} to
+find $M$-body perturbations $X_{\m w}$ that generate states of
+definite interaction energy nominally requires diagonalizing the
+matrix $\hat{\m h}$ with dimensions
 ${N\choose M}\times{N\choose M}\sim N^M\times N^M$.  If the
@@
 general prescription for exploiting the spatial symmetries of $H_0$ to
-simplify the eigenvalue problem in \eqref{eq:multi_body_eig}.
+simplify the eigenvalue problem in \eqref{eq:multi_body_eig_mat}.
 
 Our task is essentially to express the multi-body eigenvalue problem
-in \eqref{eq:multi_body_eig} in a basis that respects the spatial
+in \eqref{eq:multi_body_eig_mat} in a basis that respects the spatial
 symmetries of $H_0$.  These symmetries essentially break the set of
@@
 More generally, even if the couplings $\v{\m w}$ of $X_{\m w}$ do not
-obey the same spatial symmetries as $H_0$, these symmetries generally
-endow $\check{\m h}$ with a block-diagonal structure. \red{[TODO:
-  finish discussion/derivation]}
+obey the spatial symmetries as $H_0$, these symmetries generally endow
+$\hat{\m h}$ with a block-diagonal structure. \red{[TODO: finish
+  discussion/derivation]}
 
@@
 enforces $\bk{a|a}=1$.  We denote the set of all valid assignments of
-$N$ spins to $n$ states by $\A_n\p{N}$, such that the basis
-$\set{\ket{a}:a\in\A_n\p{N}}$ spans permutationally symmetric manifold
-of $N$ spins.  Identifying this basis allows us to expand
+$N$ (identical) spins to $n$ (distinguishable) states by $\A_n\p{N}$,
+such that the basis $\set{\ket{a}:a\in\A_n\p{N}}$ spans
+permutationally symmetric manifold of $N$ spins.  Identifying this
+basis allows us to expand
 \begin{align}

### 2020-03-22 collective_sun/sun_perturbation.tex
rewrote derivation of diagrammatic expansion of operator products

@@
 \newcommand{\I}{\mathcal{I}}
+\renewcommand{\L}{\mathcal{L}}
 \newcommand{\M}{\mathcal{M}}
 \newcommand{\N}{\mathcal{N}}
+\renewcommand{\O}{\mathcal{O}}
 \renewcommand{\P}{\mathcal{P}}
@@
 \end{align}
-where $X$ is an $M$-spin operator that is invariant under arbitrary
-permutation of its tensor factors; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is
-a choice of $M$ distinct spins that $X_k$ acts on; $\C_N\p{M}$ is the
-set of all choices of $M$ elements from a set of size $N$; and $\m w$
-is a dimension-$M$ tensor with scalar entries $w_k$.  Even though only
-${N\choose M}$ entries of $\m w$ are used to define $X_{\m w}$ in
-\eqref{eq:pert}, for later convenience we define $w_k$ for all indices
-$k\in\ZZ_N^M$ by enforcing that $\m w$ is symmetric with respect to
-any permutation of its indices, and that $w_k=0$ for any $k$ with
-repeated values. \red{[TODO: relax the restrictions on $X$ and $\m w$
-  to allow for more general perturbations, e.g.~of the form
-  $s_\x\otimes s_\y-s_\y\otimes s_\x$]}
+where $X$ is an $M$-spin operator; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is
+an ordered choice of $M$ spins that $X_k$ acts on; $\C_N\p{M}$ is the
+set of all such choices of $M$ elements from a set of size $N$; and
+$\m w$ is a dimension-$M$ tensor with scalar entries $w_k$.  Even
+though only ${N\choose M}$ entries of $\m w$ are used to define
+$X_{\m w}$ in \eqref{eq:pert}, for later convenience we define $w_k$
+for all indices $k\in\ZZ_N^M$ by enforcing that $\m w$ is symmetric
+with respect to any permutation of its indices, and that $w_k=0$ for
+any $k$ with repeated values.  In contrast, $X_k$ generally need not
+be symmetric with respect to permutations of $k$, although we will
+still define $X_k=0$ for any $k$ with repeated values.  A chiral spin
+model, for example, may have two-body interactions of the form
+$X_{pq}=s_\x^{(p)}s_\y^{(q)}-s_\y^{(p)}s_\x^{(q)}$, which are not
+invariant under exchange of spins $p$ and $q$.
 
@@
 
-Defining $\m w_\Delta$ as the projection of the tensor $\m w$ onto the
-$\Delta$-eigenspace of $\hat{\m h}$, we can expand
+Defining $\m w_\Delta$ to be the projection of the tensor $\m w$ onto
+the $\Delta$-eigenspace of $\hat{\m h}$, we can expand
 \begin{align}
@@
       \bk{m|Q_{\m v}^\dag Q_{\m v}|m}}}.
+  \label{eq:matrix_element}
 \end{align}
@@
 $m_\mu$ of the single-spin state $\mu$.
+
+\vspace{3cm}
+
+% For consistency with \eqref{eq:matrix_element}, for each state
+% $\ket{m}$ in this basis for $\M_0$ we define
+% $\ket{\1,1,m}\equiv\ket{m}$, with $\1$ the identity operator, and we
+% define $\1_1\equiv\1$.
 
@@
 \end{align}
-where $X$ is an $M$-spin operator that is invariant under arbitrary
-permutation of its tensor factors; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is
-a choice of $M$ distinct spins that $X_k$ acts on; $\C_N\p{M}$ is the
-set of all choices of $M$ elements from a set of size $N$; and $\m w$
-is a dimension-$M$ tensor with scalar entries $w_k$.  Even though only
-${N\choose M}$ entries of $\m w$ are used to define $X_{\m w}$ in
-\eqref{eq:pert_copy}, for later convenience we define $w_k$ for all
-indices $k\in\ZZ_N^M$ by enforcing that $\m w$ is symmetric with
-respect to any permutation of its indices, and that $w_k=0$ for any
-$k$ with repeated values. \red{[TODO: relax the restrictions on $X$
-  and $\m w$ to allow for more general perturbations, e.g.~of the form
-  $s_\x\otimes s_\y-s_\y\otimes s_\x$]}
+where $X$ is an $M$-spin operator; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is
+an ordered choice of $M$ spins that $X_k$ acts on; $\C_N\p{M}$ is the
+set of all such choices of $M$ elements from a set of size $N$; and
+$\m w$ is a dimension-$M$ (i.e.~$M$-index) tensor with scalar entries
+$w_k$.  Even though only ${N\choose M}$ entries of $\m w$ are used to
+define $X_{\m w}$ in \eqref{eq:pert_copy}, for later convenience we
+define $w_k$ for all indices $k\in\ZZ_N^M$ by enforcing that $\m w$ is
+symmetric with respect to any permutation of its indices, and that
+$w_k=0$ for any $k$ with repeated values.  In contrast, $X_k$
+generally need not be symmetric with respect to permutations of $k$,
+although we will still define $X_k=0$ for any $k$ with repeated
+values.
 
@@
 \label{sec:multi_body_eigenstates}
+
+\red{[TODO: fix derivation below to account for operators $X$ that are
+  not permutationally symmetric]}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Projecting products of multi-body operators onto the
-  permutationally symmetric manifold}
+\section{Permutation-symmetrized products of multi-body operators}
 \label{sec:operator_product}
@@
 Given a system of $N$ spins, we wish to project a product of
-multi-body operators onto the permutationally symmetric manifold
-$\M_0$.  For $p$ multi-body operators, the projection of such a
-product takes the form
-\begin{align}
-  \P_0 \sp{\prod_{j\in\ZZ_p}
-    \sum_{k\in\C_N\p{M_j}} w_j\p{k} X_j\p{k}} \P_0,
-  \label{eq:sym_prod_proj}
-\end{align}
-where $\P_0$ is a projector onto the PS manifold $\M_0$; $j$ indexes
-one of the $p$ multi-body operators; $X_j$ is an $M_j$-spin operator
-that is invariant under arbitrary permutation of its tensor factors;
-$k\in\C_N\p{M_j}$ is a choice of $M_j$ spins that $X_j\p{k}$ acts on;
-$\C_N\p{M_j}$ is the set of all choices of $M_j$ elements from a set
-of size $N$; and $\m w_j$ are dimension-$M_j$ tensors with scalar
-entries $w_j\p{k}$. \red{[TODO: relax the restrictions on $X_j$ and
-  $\m w_j$ to allow for more general perturbations, e.g.~of the form
-  $s_\x\otimes s_\y-s_\y\otimes s_\x$]}
-
-To simplify the product in \eqref{eq:sym_prod_proj}, we first write
-\begin{align}
-  \prod_{j\in\ZZ_p} \sum_{k\in\C_N\p{M_j}} w_j\p{k} X_j\p{k}
-  = \prod_{j\in\ZZ_p} \f1{M_j!} \sum_{k\in\ZZ_N^{M_j}}
-  w_j\p{k} X_j\p{k},
-  \label{eq:sym_prod_vec}
-\end{align}
-where we extend the definitions of $w_j\p{k}$ and $X_j\p{k}$ to all
-vectors $k\in\ZZ_N^{M_j}$ by enforcing that these objects are
-symmetric with respect to permutations of the entries in $k$, and that
-$w_j\p{k}=X_j\p{k}=0$ if $k$ contains any repeated entries.  We then
-collect terms whose operator content is equivalent up to a permutation
-of spins.  To this end, we classify terms in \eqref{eq:sym_prod_vec}
-by the numbers $g_S$ of indices shared by all tensors $w_j$ with
-$j\in S\subset\ZZ_p$.  For example, a term of the form
+multi-body operators onto the permutationally symmetric manifold (PS)
+$\M_0$.  For $p$ multi-body operators, such a product takes the form
+\begin{align}
+  \m X_{\m w}
+  \equiv \prod_{\alpha\in\ZZ_p} \sum_{k_\alpha\in\C_N\p{M_\alpha}}
+  w_\alpha\p{k_\alpha} X_\alpha\p{k_\alpha},
+  \label{eq:sym_prod_start}
+\end{align}
+where $\m X\equiv\p{X_1,X_2,\cdots,X_p}$ is a vector of
+$M_\alpha$-spin operators $X_\alpha$;
+$\m w\equiv\p{w_1,w_2,\cdots,w_p}$ is a vector of dimension-$M_\alpha$
+tensors $w_\alpha$;
+$k_\alpha\equiv\p{k_{\alpha,1},k_{\alpha,2},\cdots,k_{\alpha,M_\alpha}}$
+is an ordered choice of $M_\alpha$ spins that $X_\alpha\p{k_\alpha}$
+acts on; $\C_N\p{M_\alpha}$ is the set of all such choices of
+$M_\alpha$ elements from a set of size $N$; and $w_\alpha\p{k_\alpha}$
+is a scalar entry of $w_\alpha$.  Although only ${N \choose M_\alpha}$
+entries of $w_\alpha$ are used in \eqref{eq:sym_prod_start}, we define
+$w_\alpha\p{k_\alpha}$ for all $k_\alpha\in\ZZ_N^{M_\alpha}$ by
+enforcing that $w_\alpha\p{k_\alpha}$ is symmetric with respect to
+permutations of $k_\alpha$, and that $w_\alpha\p{k_\alpha}=0$ if
+$k_\alpha$ has any repeated values.  In contrast,
+$X_\alpha\p{k_\alpha}$ generally need not be symmetric with respect to
+permutations of $k_\alpha$, although we still enforce that
+$X_\alpha\p{k_\alpha}=0$ if $k_\alpha$ has any repeated values.
+
+To simplify the product in \eqref{eq:sym_prod_start}, we rewrite the
+product of sums as a sum of products, and symmetrize over all
+permutations of $k_\alpha\in\C_N\p{M_\alpha}$ in order to make all
+individual sums independent:
+\begin{align}
+  \m X_{\m w}
+  &= \f1{M_1!} \sum_{k_1\in\ZZ_N^{M_1}}
+  \f1{M_2!} \sum_{k_2\in\ZZ_N^{M_2}} \cdots
+  \f1{M_p!} \sum_{k_p\in\ZZ_N^{M_p}}
+  \prod_{\alpha\in\ZZ_p} w_\alpha\p{k_\alpha} X_\alpha\p{k_\alpha} \\
+  &= \f1{\m M!} \sum_{k\in\ZZ_N^{\m M}} \prod_{\alpha\in\ZZ_p}
+  w_\alpha\p{k_\alpha} X_\alpha\p{k_\alpha},
+  \label{eq:sum_of_products}
+\end{align}
+where $\m M\equiv\p{M_1,M_2,\cdots,M_p}$ is a vector of the
+dimensions $M_\alpha$; $\m M!\equiv\prod_{M_\alpha\in\m M}M_\alpha!$
+for shorthand; and
+$\ZZ_N^{\m M}\equiv\bigotimes_{\alpha\in\ZZ_p}\ZZ_N^{M_\alpha}$, such
+that each $k\in\ZZ_N^{\m M}$ has the decomposition
+$k=\p{k_1,k_2,\cdots,k_p}$ with $k_\alpha\in\ZZ_N^{M_\alpha}$.
+
+We now classify terms in \eqref{eq:sum_of_products} by the numbers
+$g_S$ of indices shared by all tensors $w_\alpha$ with
+$\alpha\in S\subset\ZZ_p$.  For example, a term of the form
 $w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}$ with distinct indices
-$\p{a,b,c,d,e}\in\C_N\p{5}$ would have
+$a,b,c,d,e$ would have
 \begin{align}
@@
 and $g_S=0$ for all other subsets $S\subset\ZZ_3$.  This index
-assignment can be represented by the Venn diagram
+assignment can be associated with the Venn diagram
 \begin{align}
@@
 where $g_S$ is determined the number of dots at the intersection of
-circles $\set{w_j:j\in S}$.  Denoting the set of all nonempty subsets
-of $\ZZ_p$ by $\PP^*\p{\ZZ_p}$, we denote an assignment of distinct
-indices according to choices of $g_S$ for all $S\in\PP^*\p{\ZZ_p}$ by
-$g$.  Given a vector $\m M$ of the dimensions $M_j$, we then denote
-the set of all valid index assignments $g$ by $\G\p{\m M}$.  For
-consistency, all valid index assignments $g\in\G\p{\m M}$ must assign
-exactly $M_j$ indices to the dimension-$M_j$ tensor $w_j$, i.e.~
-\begin{align}
-  \sum_{S\in\PP^*\p{\ZZ_p}\,:\,j\in S} g_S
-  = \sum_{R\in\PP^*\p{\ZZ_p\setminus\set{j}}} g_{\set{j}\cup R}
-  = M_j
-\end{align}
-for all $j\in\ZZ_p$.  The set of valid index assignments $\G\p{\m M}$
-is then essentially the set of all Venn diagrams of the form in
-\eqref{eq:venn_diagram} with $p=\dim\p{\m M}$ circles and a total of
-$M_j$ dots within circle $w_j$.
-
-A classification of terms in \eqref{eq:sym_prod_vec} by index
-assignments $g\in\G\p{\m M}$ allows us to expand
-\begin{align}
-  \prod_{j\in\ZZ_p} \sum_{k\in\ZZ_N^{M_j}} w_j\p{k} X_j\p{k}
-  \EQPS \sum_{g\in\G\p{\m M}} \S\p{g} w\p{g} X\p{g},
-  \label{eq:sym_prod_group_start}
-\end{align}
-where $\EQPS$ denotes equality up to a restriction to the
-permutationally symmetric manifold; $X\p{g}$ is an operator acquired
-by assigning indices (choices of spins) to all $X_j$ in a manner
-consistent with $g$; $w\p{g}$ is a scalar acquired by summing over all
-indices assigned to the tensors $w_j$ according to $g$; and $\S\p{g}$
-is a symmetry factor accounting for the number of ways to assign
-indices according to $g$.
-
-In order to write out the factors in \eqref{eq:sym_prod_group_start}
-explicitly, we identify the set of values that the distinct indices
-assigned according to $g$ can take:
-\begin{align}
-  \C_N\p{g}
-  \equiv \set{ k \in \bigotimes_{S\in\PP^*\p{\ZZ_p}} \C_N\p{g_S}
-    : \t{all elements of $k$ are distinct} },
-  \label{eq:index_values}
-\end{align}
-where each $k\in\C_N\p{g}$ has the natural decomposition
-$k=\p{k_S:S\in\PP^*\p{\ZZ_p}}$ with $k_S\in\C_N\p{g_S}$.  In words,
-$\C_N\p{g}$ consists of all the ways to choose, for each subset $S$ of
-the tensors $\set{w_j}$, a set $k_S$ of $g_S$ distinct spins, with the
-restriction that all $k_S$ are orthogonal: if $z\in k_S$, then
-$z\notin k_R$ for $R\ne S$.  Having defined the set $\C_N\p{g}$ of
-index values, we define the restriction of $k\in\C_N\p{g}$ to values
-associated with $w_j$ and $X_j$:
-\begin{align}
-  k_j \equiv \bigcup_{S\in\PP^*\p{\ZZ_p} : j\in S} k_S
-  = \bigcup_{R\in\PP^*\p{\ZZ_p\setminus\set{j}}} k_{\set{j}\cup R},
-\end{align}
-where the union of lists $k_S$ denotes a concatenation of those lists.
-These definitions allow us to expand
-\begin{align}
-  w\p{g} \equiv \sum_{k\in\C_N\p{g}} \prod_{j\in\ZZ_p} w_j\p{k_j},
-  &&
-  X\p{g} \equiv \prod_{j\in\ZZ_p}
-  X_j\p{\ell_j}~\t{for some}~\ell\in\C_N\p{g},
-  \label{eq:diagram_factors}
-\end{align}
-where the choice of $\ell\in\C_N\p{g}$ does not matter after
-projection onto permutationally symmetric manifold.  Finally, we can
-work out that
-\begin{align}
-  \S\p{g} \equiv \sp{\prod_{j\in\ZZ_p} \C\p{ g_S : j\in S }}
-  \sp{\prod_{S\in\PP^*\p{\ZZ_p}} \p{g_S!}^\abs{S}}
-  = \prod_{j\in\ZZ_p} M_j!,
-  \label{eq:sym_prod_symmetry}
-\end{align}
-where the first product in \eqref{eq:sym_prod_symmetry} contains
+circles $\alpha\in S$.  We denote the set of all nonempty subsets of
+$\ZZ_p$ by $\PP^*\p{\ZZ_p}$, and denote a choice of $g_S$ for all
+$S\in\PP^*\p{\ZZ_p}$ by $g$.  With representations such as
+\eqref{eq:venn_diagram} in mind, we will loosely refer to $g$ as
+simply a ``diagram''.  We define
+$\abs{g}\equiv\sum_{S\in\PP^*\p{\ZZ_p}}g_S$ to be the number of
+indices considered in $g$ (or equivalently the number of dots in the
+diagram $g$), and we denote the set of all diagrams for
+$p\equiv\dim\p{\m M}$ tensors with dimensions $\m M$ by $\G\p{\m M}$.
+For consistency, all diagrams $g\in\G\p{\m M}$ must associate exactly
+$M_\alpha$ indices with the dimension-$M_\alpha$ tensor $w_\alpha$,
+i.e.
+\begin{align}
+  \sum_{S\in\PP^*\p{\ZZ_p}\,:\,\alpha\in S} g_S
+  = \sum_{R\in\PP^*\p{\ZZ_p\setminus\set{\alpha}}} g_{\set{\alpha}\cup R}
+  = M_\alpha
+\end{align}
+for all $\alpha\in\ZZ_p$.  Schematically, a classification of the
+terms in \eqref{eq:sum_of_products} by diagrams $g\in\G\p{\m M}$ will
+allow us to expand
+\begin{align}
+  \m X_{\m w} \EQPS \sum_{g\in\G\p{\m M}} \m w\p{g} \m X\p{g},
+  \label{eq:sym_prod_teaser}
+\end{align}
+where $\EQPS$ denotes equality up to a restriction to the PS manifold
+$\M_0$; $\m w\p{g}$ is a scalar acquired by assigning $\abs{g}$
+indices to $\m w$ according to $g$, and summing over those indices in
+some appropriate manner; and $\m X\p{g}$ is a $\abs{g}$-spin operator
+acquired by averaging over all of the distinct ways to assign
+$\abs{g}$ indices to $\m X$ according to $g$ (clarified below).
+
+In order to derive the expansion in \eqref{eq:sym_prod_teaser} and
+find the explicit forms of $\m w\p{g}$ and $\m X\p{g}$, we need to
+collect all terms in \eqref{eq:sum_of_products} that are associated
+with the diagram $g$.  To this end, we first consider all distinct
+ways to assign $\abs{g}$ indices to specific tensor factors of $\m w$
+and $\m X$ in a manner consistent with $g$.  For example, if
+$\m X=\p{X_1,X_2}$ with dimensions $\p{M_1,M_2}=\p{2,1}$ and
+$\abs{g}=2$ with $g_{\set{1,2}}=g_{\set{2}}=1$, then
+$X_1\p{a,b} X_2\p{b}$ and $X_1\p{b,a} X_2\p{b}$ are distinct
+assignments of $\abs{g}$ indices to $\m X$ in a manner consistent with
+$g$, provided that $a\ne b$.  After assigning $\abs{g}$ indices to
+$\m w$ and $\m X$ in a manner consistent with $g$, one needs to sum
+over all values that these indices can take, which amounts to a sum
+over all $k=\p{k_1,k_2,\cdots,k_{\abs{g}}}\in\ZZ_N^{\abs{g}}$ with the
+restriction that $k_\mu\ne k_\nu$ for all $\mu\ne\nu$.
+
+A specific assignment of indices can be thought of as a choice, for
+each dot in a region $S$ of a diagram $g$, of which tensor factor of
+$w_\alpha$ and $X_\alpha$ (for all $\alpha\in S$) that dot is
+associated with.  The only catch is that such choices treat all dots
+in $g$ as distinct, whereas the $g_S$ dots within region $S$ of $g$
+are actually indistinguishable.  The assignment
+$X_1\p{a,b,c} X_2\p{a,b}$, for example, is the same as
+$X_1\p{b,a,c} X_2\p{b,a}$.  We denote the set of all distinct index
+assignments according to $g$ by $\L\p{g}$.  The number of such
+assignments is
+\begin{align}
+  \abs{\L\p{g}}
+  = \sp{\prod_{\alpha\in\ZZ_p} \C\p{ g_S : \alpha\in S }}
+  \sp{\prod_{S\in\PP^*\p{\ZZ_p}} \p{g_S!}^{\abs{S}}}
+  \sp{\prod_{S\in\PP^*\p{\ZZ_p}} g_S!}^{-1}
+  = \m M! \sp{\prod_{S\in\PP^*\p{\ZZ_p}} g_S!}^{-1},
+  \label{eq:index_assignment_num}
+\end{align}
+where the first product in \eqref{eq:index_assignment_num} contains
 multinomial coefficients
@@
 \end{align}
-that account for the number of ways to partition the indices of each
-$w_j$ into sets of shared indices as specified by $g$, and the second
-product in \eqref{eq:sym_prod_symmetry} accounts for the number of
-ways to permute each set of shared indices on every tensor.
-Altogether, the symmetry factors $\S\p{g}$ cancel out with the
-$\sim M_j!$ prefactors in \eqref{eq:sym_prod_vec}, so
-\begin{align}
-  \prod_{j\in\ZZ_p} \sum_{k\in\C_N\p{M_j}} w_j\p{k} X_j\p{k}
-  \EQPS \sum_{g\in\G\p{\m M}} w\p{g} X\p{g}.
-  \label{eq:sym_prod_group}
-\end{align}
-This result essentially gives us a diagrammatic expansion for the
-product of multi-body operators, projected onto the permutationally
-symmetric manifold.
+that account for the number of ways to partition the $M_\alpha$
+indices of each $w_\alpha$ and $X_\alpha$ into sets of shared indices
+as specified by $g$, the second product in
+\eqref{eq:index_assignment_num} accounts for the number of ways to
+permute each set of shared indices on each $w_\alpha$ and $X_\alpha$,
+and the third product in \eqref{eq:index_assignment_num} accounts for
+the treatment of the dots within each region $S$ of $g$ as
+distinguishable, when in fact they are not.  Altogether, we have
+\begin{align}
+  \m X_{\m w}
+  = \f1{\m M!} \sum_{g\in\G\p{\m M}} \sum_{\ell\in\L\p{g}}
+  \sum_{\substack{k\in\ZZ_N^{\abs{g}}\\
+      k_\mu\ne k_\nu\forall \mu\ne\nu}}
+  \m w\p{k,\ell} \m X\p{k,\ell},
+\end{align}
+where $\m w\p{k,\ell}$ and $\m X\p{k,\ell}$ denote the assignment of
+indices $k$ to all $w_\alpha$ and $X_\alpha$ according to $\ell$.  The
+invariance of each tensor $w_\alpha$ under arbitrary permutations of
+its indices means that all assignments of $\abs{g}$ indices to $\m w$
+consistent with $g$ are equivalent,
+e.g.~$w_1\p{a,b} w_2\p{b} = w_1\p{b,a} w_2\p{b}$.  The value of
+$\m w\p{k,\ell}$ therefore depends on $g$, but not the specific choice
+of $\ell\in\L\p{g}$.  Furthermore, the specific choice of spins $k$ in
+$\m X\p{k,\ell}$ does not matter after projection onto the PS manifold
+$\M_0$, so the dependence of $\m X\p{k,\ell}$ on $k$ is similarly
+superfluous, which is to say that we can write
+\begin{align}
+  \m X_{\m w}
+  \EQPS \f1{\m M!} \sum_{g\in\G\p{\m M}}
+  \sum_{\substack{k\in\ZZ_N^{\abs{g}}\\
+      k_\mu\ne k_\nu\forall \mu\ne\nu}}
+  \m w\p{k,g} \sum_{\ell\in\L\p{g}} \m X\p{\ell}.
+\end{align}
+Defining the average of $\m X\p{\ell}$ over all $\ell\in\L\p{g}$,
+\begin{align}
+  \m X\p{g} \equiv \f1{\abs{\L\p{g}}}
+  \sum_{\ell\in\L\p{g}} \m X\p{\ell},
+\end{align}
+we can more simply write
+\begin{align}
+  \m X_{\m w}
+  \EQPS \f1{\m M!} \sum_{g\in\G\p{\m M}} \abs{\L\p{g}}
+  \sum_{\substack{k\in\ZZ_N^{\abs{g}}\\
+      k_\mu\ne k_\nu\forall \mu\ne\nu}}
+  \m w\p{k,g} \m X\p{g}.
+\end{align}
+We now examine the sum over all $k\in\ZZ_N^{\abs{g}}$ with
+$k_\mu\ne k_\nu\forall \mu\ne\nu$.  For any given instance of
+$\m w\p{k,g}$, each entry $k_\mu\in k$ is associated with tensors
+$\set{w_\alpha:\alpha\in S}$ for some subset $S\in\PP^*\p{\ZZ_p}$.
+Identifying the indices associated with $S\in\PP^*\p{\ZZ_p}$ by
+$k_S\in\ZZ_N^{g_S}$ and enforcing the decomposition
+$k=\p{k_S:S\in\PP^*\p{\ZZ_p}}$, we can write
+\begin{align}
+  \sum_{\substack{k\in\ZZ_N^{\abs{g}}\\
+      k_\mu\ne k_\nu\forall \mu\ne\nu}} \m w\p{k,g}
+  = \sum_{k\in\ZZ_N\p{g}} \m w\p{k},
+\end{align}
+where
+\begin{align}
+  \ZZ_N\p{g} \equiv
+  \set{ k \in \bigotimes_{S\in\PP^*\p{\ZZ_p}} \ZZ_N^{g_S}
+    : \t{all elements of $k$ are distinct} }.
+\end{align}
+The domain $\ZZ_N\p{g}$ is redundant, however, in the sense that
+$\m w\p{k}$ for any $k\in\ZZ_N\p{g}$ is invariant under permutations
+of each $k_S\in\ZZ_N\p{g_S}$, so
+\begin{align}
+  \sum_{k\in\ZZ_N\p{g}} \m w\p{k}
+  = \sp{\prod_{S\in\PP^*\p{\ZZ_p}}g_S!} \sum_{k\in\C_N\p{g}} \m w\p{k},
+\end{align}
+where
+\begin{align}
+  \C_N\p{g} \equiv
+  \set{ k \in \bigotimes_{S\in\PP^*\p{\ZZ_p}} \C_N\p{g_S}
+    : \t{all elements of $k$ are distinct} }.
+\end{align}
+Defining
+\begin{align}
+  \m w\p{g} \equiv \sum_{k\in\C_N\p{g}} \m w\p{k},
+  \label{eq:diagram_factors}
+\end{align}
+we thus find that
+\begin{align}
+  \m X_{\m w} \EQPS \sum_{g\in\G\p{\m M}} \m w\p{g} \m X\p{g},
+  \label{eq:sym_prod_diagrams}
+\end{align}
+where all combinatorial factors such as $\m M!$ and $\abs{\L\p{g}}$
+cancel each other out.
 
@@
 Diagrams such as \eqref{eq:venn_diagram} will play a major role in the
-calculations in this section.  Here, we define these diagrams as
-formal objects that we can use in mathematical expressions, enabling
-us to systematically perform calculations that are otherwise
-intractable.  We then derive rules to ``simplify'' these diagrams in
-such a way as to reduce the computational cost of their numerical
-evaluation.
+expansion in \eqref{eq:sym_prod_diagrams}, as well as the remainder of
+this appendix.  Here, we define these diagrams as formal objects that
+we can use in mathematical expressions, enabling us to systematically
+perform calculations that are otherwise intractable.  We then derive
+rules to ``simplify'' these diagrams in such a way as to reduce the
+computational cost of their numerical evaluation.
 
@@
 intersection of the circles $j\in S$, and are simply equal to the
-associated scalar $w\p{g}$ defined in \eqref{eq:diagram_factors}, e.g.
+associated scalar $\m w\p{g}$ defined in \eqref{eq:diagram_factors},
+e.g.
 \begin{align}
@@
 The process of constructing and simplifying diagrams to compute the
-coefficients $w\p{g}$ of the diagrammatic operator product expansion
-in \eqref{eq:sym_prod_group} is systematic enough to be converted into
-an algorithm for execution on a computer.  The last step necessary to
-automate the evaluation of \eqref{eq:sym_prod_group} is expansion of
-the multi-spin operators $X\p{g}$ in a basis for the permutationally
-symmetric manifold $\M_0$.  Without loss of generality, we consider a
-single permutationally symmetric $M$-spin operator $X$; our task is
-essentially to find coefficients of the expansion
+coefficients $\m w\p{g}$ of the diagrammatic operator product
+expansion in \eqref{eq:sym_prod_diagrams} is systematic enough to be
+converted into an algorithm for execution on a computer.  The last
+step necessary to automate the evaluation of
+\eqref{eq:sym_prod_diagrams} is expansion of the multi-spin operators
+$\m X\p{g}$ in a basis for the PS manifold $\M_0$.  Without loss of
+generality, we consider a single permutationally symmetric $M$-spin
+operator $X$; our task is essentially to find coefficients of the
+expansion
 \begin{align}
@@
 \begin{align}
-  \P_0 X_{\m v} Y_{\m w} \P_0
-  = \P_0 \sp{\sum_{p,q\in\ZZ_N} v_p w_q X_p Y_q} \P_0
+  X_{\m v} Y_{\m w}
+  = \sp{\sum_{p,q\in\ZZ_N} v_p w_q X_p Y_q}
   \EQPS \diagram{single_body_0} X_1 Y_2
@@
 \end{align}
-where $\P_0$ is a projector onto the permutationally symmetric
-manifold $\M_0$; $p,q$ index individual spins; $v_p,w_p$ are scalars;
-and $X,Y$ are single-spin operators.  This product appears, for
-example, in the calculation of the second-order effective Hamiltonian
-$H_1^{(2)}$ induced on the permutationally symmetric manifold $\M_0$
-by a single-body perturbation.  Defining
+where $\P_0$ is a projector onto the PS manifold $\M_0$; $p,q$ index
+individual spins; $v_p,w_p$ are scalars; and $X,Y$ are single-spin
+operators.  This product appears, for example, in the calculation of
+the second-order effective Hamiltonian $H_1^{(2)}$ induced on the
+permutationally symmetric manifold $\M_0$ by a single-body
+perturbation.  Defining
 \begin{align}
@@
 \begin{multline}
-  \P_0 X_{\m v} Y_{\m w} \P_0
-  = \P_0 \sp{\sum_{\p{k,\ell},\p{p,q}\in\C_N\p{2}}
-    v_{k\ell} w_{pq} X_{k\ell} Y_{pq}} \P_0 \\
+  X_{\m v} Y_{\m w}
+  = \sp{\sum_{\p{k,\ell},\p{p,q}\in\C_N\p{2}}
+    v_{k\ell} w_{pq} X_{k\ell} Y_{pq}} \\
   \EQPS \diagram{two_body_0} X_{1,2} Y_{3,4}
-  + \diagram{two_body_1} X_{1,2} Y_{1,3}
-  + \diagram{two_body_2} X_{1,2} Y_{1,2},
+  + \diagram{two_body_1} \sp{X,Y}_{1,2,3}
+  + \diagram{two_body_2} \sp{X,Y}_{1,2},
 \end{multline}
-where $\P_0$ is a projector onto the permutationally symmetric
-manifold $\M_0$; $k,\ell,p,q$ index individual spins;
-$v_{k\ell},w_{pq}$ are scalar coefficients with $m_{pq}=m_{qp}$ and
-$m_{pp}=0$ for both $\m m\in\set{\m v,\m w}$; and $X,Y$ are
-permutationally symmetric two-spin operators.  This product appears,
-for example, in the calculation of the second-order effective
-Hamiltonian $H_2^{(2)}$ induced on the permutationally symmetric
-manifold $\M_0$ by a two-body perturbation.  Defining
+where $\P_0$ is a projector onto the PS manifold $\M_0$; $k,\ell,p,q$
+index individual spins; $v_{k\ell},w_{pq}$ are scalar coefficients
+with $m_{pq}=m_{qp}$ and $m_{pp}=0$ for both $\m m\in\set{\m v,\m w}$;
+and $X,Y$ are permutationally symmetric two-spin operators; and
+\begin{align}
+  \sp{X,Y}_{1,2,3}
+  \equiv \f14 X_{1,2}\p{Y_{1,3} + Y_{2,3} + Y_{3,1} + Y_{3,2}},
+  &&
+  \sp{X,Y}_{1,2} \equiv \f12 X_{1,2} \p{Y_{1,2} + Y_{2,1}}.
+\end{align}
+This product appears, for example, in the calculation of the
+second-order effective Hamiltonian $H_2^{(2)}$ induced on the
+permutationally symmetric manifold $\M_0$ by a two-body perturbation.
+Defining
 \begin{align}
@@
   X_{1,2} Y_{3,4}
-  + \p{\v v\c\v w - 2 \v{\m v} \c \v{\m w}} X_{1,2} Y_{1,3}
-  + \v{\m v} \c \v{\m w}\, X_{1,2} Y_{1,2}.
+  + \p{\v v\c\v w - 2 \v{\m v} \c \v{\m w}} \sp{X,Y}_{1,2,3}
+  + \v{\m v} \c \v{\m w}\, \sp{X,Y}_{1,2}.
 \end{align}
@@
 \begin{align}
-  \Z \equiv \P_0 \sp{\sum_{\p{k,\ell},\p{p,q},\p{r,s}\in\C_N\p{2}}
-    u_{k\ell} v_{pq} w_{rs} Z_k Z_\ell Z_p Z_q Z_r Z_s} \P_0,
+  Z_{\m u\m v\m w}
+  \equiv \sum_{\p{k,\ell},\p{p,q},\p{r,s}\in\C_N\p{2}}
+    u_{k\ell} v_{pq} w_{rs} Z_k Z_\ell Z_p Z_q Z_r Z_s,
   &&
@@
 \end{align}
-where $\P_0$ is a projector onto the permutationally symmetric
-manifold $\M_0$; $k,\ell,p,q,r,s$ index individual spins; $Z$ is a
-single-spin operator; and $u_{k\ell},v_{pq},w_{rs}$ are scalar
-coefficients with $m_{pq}=m_{qp}$ and $m_{pp}=0$ for each of
-$\m m\in\set{\m u,\m v,\m w}$.  \red{[TODO: give example of when this
-  product appears.]} Collecting terms according to the number of $Z$
+projected onto the PS manifold $\M_0$, where $k,\ell,p,q,r,s$ index
+individual spins; $Z$ is a single-spin operator; and
+$u_{k\ell},v_{pq},w_{rs}$ are scalar coefficients with $m_{pq}=m_{qp}$
+and $m_{pp}=0$ for each of $\m m\in\set{\m u,\m v,\m w}$.  This
+product appears, for example, in the projection of Ising-type
+($s_\z\otimes s_\z$) interactions onto the two-body excitation
+manifold $\M_2$.  Collecting terms according to the number of $Z$
 operators that remain after accounting for $Z^2=1$, we find that
 \begin{align}
-  \Z \EQPS A_6 Z^{\otimes 6} + A_4 Z^{\otimes 4}
+  Z_{\m u\m v\m w} \EQPS A_6 Z^{\otimes 6} + A_4 Z^{\otimes 4}
   + A_2 Z^{\otimes 2} + A_0,
@@
 \end{multline}
-Finally, we can also write the product $\Z$ in \eqref{eq:triple_multi}
-in terms of the collective operator
+Finally, we can also write the product $Z_{\m u\m v\m w}$ in
+\eqref{eq:triple_multi} in terms of the collective operator
 $\col{Z} \equiv \sum_{p\in\ZZ_N} Z_p$:
 \begin{align}
-  \Z \EQPS
+  Z_{\m u\m v\m w} \EQPS
   \tilde A_6 \col{Z}^6 + \tilde A_4 \col{Z}^4

### 2020-03-22 collective_sun/sun_perturbation.tex
fixed derivation of multi-body eigenvalue problem to account for operators that are not permutationally symmetric

@@
 
-\red{[TODO: fix derivation below to account for operators $X$ that are
-  not permutationally symmetric]}
+\red{[TODO: verify that the derivation below properly accounts for
+  operators $X$ that are not permutationally symmetric]}
 
@@
 permutation operator $\Pi_{pq}$ acts trivially on $X_k\ket\psi$ when
-$p,q\in k$ or $p,q\notin k$, leaving terms at each fixed $k$ of the
-form
+$p,q\notin k$.  When $p,q\in k$, this permutation operator swaps the
+positions of $p$ and $q$ in $X_k$, but we can simply relabel
+$\p{p,q}\leftrightarrow\p{q,p}$ with no consequence for $w_k$, so
+\begin{align}
+  \sum_{\substack{k\in\C_N\p{M}\\\p{p,q}\in\C_N\p{2}\\
+      p,q\notin k~\t{or}~p,q\in k}}
+  h_{pq} w_k \Pi_{pq} X_k \ket\psi
+  = \sum_{k\in\C_N\p{M}}
+  \sum_{\substack{\p{p,q}\in\C_N\p{2}\\p,q\notin k~\t{or}~p,q\in k}}
+  h_{pq} w_k X_k \ket\psi.
+\end{align}
+For fixed $k\in\C_N\p{M}$, we can simplify
 \begin{align}
@@
   = \sum_{\p{p,q}\in\C_N\p{2}} h_{pq}
-  - \sum_{\substack{p\in k\\q\notin k}} h_{pq}
+  - \sum_{\substack{p\in k\\q\in\ZZ_N\setminus k}} h_{pq}
   = E_0 - \tilde h_k,
@@
 that $p,q\in k$ or $p,q\notin k$ into an unrestricted sum minus a
-remainder with the complementary restriction.  The terms in
+remainder with the complementary restriction; and $E_0$ is the
+interaction energy of PS states $\ket\psi\in\M_0$.  The terms in
 \eqref{eq:diagnosis_start} with $p,q\in k$ or $p,q\notin k$ are then

### 2020-03-22 collective_sun/sun_perturbation.tex
simplified explanation/counting of different index assignments

@@
 A specific assignment of indices can be thought of as a choice, for
-each dot in a region $S$ of a diagram $g$, of which tensor factor of
-$w_\alpha$ and $X_\alpha$ (for all $\alpha\in S$) that dot is
-associated with.  The only catch is that such choices treat all dots
-in $g$ as distinct, whereas the $g_S$ dots within region $S$ of $g$
-are actually indistinguishable.  The assignment
-$X_1\p{a,b,c} X_2\p{a,b}$, for example, is the same as
-$X_1\p{b,a,c} X_2\p{b,a}$.  We denote the set of all distinct index
-assignments according to $g$ by $\L\p{g}$.  The number of such
-assignments is
-\begin{align}
-  \abs{\L\p{g}}
-  = \sp{\prod_{\alpha\in\ZZ_p} \C\p{ g_S : \alpha\in S }}
-  \sp{\prod_{S\in\PP^*\p{\ZZ_p}} \p{g_S!}^{\abs{S}}}
-  \sp{\prod_{S\in\PP^*\p{\ZZ_p}} g_S!}^{-1}
-  = \m M! \sp{\prod_{S\in\PP^*\p{\ZZ_p}} g_S!}^{-1},
-  \label{eq:index_assignment_num}
-\end{align}
-where the first product in \eqref{eq:index_assignment_num} contains
-multinomial coefficients
-\begin{align}
-  \C\p{a_1,a_2,\cdots,a_m}
-  \equiv \f{\p{a_1+a_2+\cdots+a_m}!}{a_1!a_2!\cdots a_m!}
-\end{align}
-that account for the number of ways to partition the $M_\alpha$
-indices of each $w_\alpha$ and $X_\alpha$ into sets of shared indices
-as specified by $g$, the second product in
-\eqref{eq:index_assignment_num} accounts for the number of ways to
-permute each set of shared indices on each $w_\alpha$ and $X_\alpha$,
-and the third product in \eqref{eq:index_assignment_num} accounts for
-the treatment of the dots within each region $S$ of $g$ as
-distinguishable, when in fact they are not.  Altogether, we have
+each of the $M_\alpha$ tensor factors of $w_\alpha$ and $X_\alpha$, of
+which dot in the diagram $g$ that tensor factor is associated with
+(i.e.~out of the $M_\alpha$ dots within circle $\alpha$).  There are
+nominally $M_\alpha!$ such choices for a given $\alpha$, and therefore
+$\prod_{\alpha\in\ZZ_p}M_\alpha!=\m M!$ choices total.  The only catch
+is that these choices treat all dots in $g$ as distinct, whereas the
+$g_S$ dots within region $S$ of $g$ are actually indistinguishable.
+The assignment $X_1\p{a,b,c} X_2\p{a,b}$, for example, is the same as
+$X_1\p{b,a,c} X_2\p{b,a}$.  In general, the treatment of the $g_S$
+indistinguishable dots within region $S\in\PP^*\p{\ZZ_p}$ of $g$ as
+distinguishable results in over-counting distinct index assignments by
+a factor of $g_S!$.  Denoting the set of all distinct index
+assignments according to $g$ by $\L\p{g}$, the number of such
+assignments is therefore
+\begin{align}
+  \abs{\L\p{g}} = \m M! \sp{\prod_{S\in\PP^*\p{\ZZ_p}} g_S!}^{-1}.
+\end{align}
+Altogether, we have
 \begin{align}

### 2020-03-24 collective_sun/sun_perturbation.tex
major changes allowing for the consideration of more general multi-body operators

@@
 \newcommand{\C}{\mathcal{C}}
+\newcommand{\D}{\mathcal{D}}
 \newcommand{\E}{\mathcal{E}}
@@
 \DeclareMathOperator{\var}{var}
+\DeclareMathOperator{\cov}{cov}
 \DeclareMathOperator{\diag}{diag}
@@
 interaction Hamiltonian $H_0$ in \eqref{eq:H_0}, we will add to the
-Hamiltonian an operator of the general form \red{[TODO: consider more
-  general sums of such operators]}
-\begin{align}
-  X_{\m w} \equiv \sum_{k\in\C_N\p{M}} w_k X_k,
+Hamiltonian an operator $\hat\O$ of the general form
+\begin{align}
+  \hat\O \equiv \sum_{\p{X,\m w}\in\O} X_{\m w},
+  &&
+  X_{\m w} \equiv \f1{M!} \sum_{k\in\D_N\p{M}} w_k X_k,
   \label{eq:pert}
 \end{align}
-where $X$ is an $M$-spin operator; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is
-an ordered choice of $M$ spins that $X_k$ acts on; $\C_N\p{M}$ is the
-set of all such choices of $M$ elements from a set of size $N$; and
-$\m w$ is a dimension-$M$ tensor with scalar entries $w_k$.  Even
-though only ${N\choose M}$ entries of $\m w$ are used to define
-$X_{\m w}$ in \eqref{eq:pert}, for later convenience we define $w_k$
-for all indices $k\in\ZZ_N^M$ by enforcing that $\m w$ is symmetric
-with respect to any permutation of its indices, and that $w_k=0$ for
-any $k$ with repeated values.  In contrast, $X_k$ generally need not
-be symmetric with respect to permutations of $k$, although we will
-still define $X_k=0$ for any $k$ with repeated values.  A chiral spin
-model, for example, may have two-body interactions of the form
-$X_{pq}=s_\x^{(p)}s_\y^{(q)}-s_\y^{(p)}s_\x^{(q)}$, which are not
-invariant under exchange of spins $p$ and $q$.
+where $\hat\O$ is constructed from the set $\O\equiv\set{\p{X,\m w}}$
+of $M$-spin operators $X$ together with dimension-$M$ (i.e.~$M$-index)
+tensors $\m w$; $k=\p{k_1,k_2,\cdots,k_M}$ is a list of of $M$ spins
+that $X_k$ acts on; $w_k$ is a scalar entry of $\m w$; and
+\begin{align}
+  \D_N\p{M} \equiv
+  \set{ k \in \ZZ_N^M:~\t{all elements of}~k~\t{are distinct} },
+  \label{eq:off_diags}
+\end{align}
+is the set of all ordered choices of $M$ spins out of $N$.  The
+normalization factor of $1/M!$ in \eqref{eq:pert} is included to
+account for the fact that there are $M!$ terms addressing any given
+choice of $M$ of spins.  Due to the permutational symmetry of the
+ground-state manifold $\M_0$ of the interaction Hamiltonian $H_0$, we
+will find it convenient to decompose
+$\D_N\p{M}=\pi\p{M}\times\C_N\p{M}$, where $\pi\p{M}$ is the
+permutation group of $\ZZ_M$, and $\C_N\p{M}$ is the set of all
+choices of $M$ spins out of $N$.  We can then expand
+\begin{align}
+  X_{\m w} = \f1{M!} \sum_{\sigma\in\pi\p{M}} \sum_{k\in\C_N\p{M}}
+  w_{\sigma\p{k}} X_{\sigma\p{k}}.
+\end{align}
+For simplicity, we will assume throughout this work that $\O$ is
+homogeneous in the sense that $M$ is the same for all
+$\p{X,\m w}\in\O$, with the understanding that a generalization to
+different $M$ for each $\p{X,\m w}\in\O$ is straightforward.
 
@@
 considering this case will be necessary to construct low-lying
-eigenstates of $H_0$.  Constructing low-lying eigenstates of $H_0$
-will in turn enable simulating the dynamics induced by
-non-perturbative additions $X_{\m w}$ to the SU($n$) Heisenberg model,
-with norms $\norm{X_{\m w}}\sim\Delta_{\t{gap}}$.
-
-When $X_{\m w}$ is sufficiently small, we can treat its effect on the
-PS manifold $\M_0$ perturbatively.  The effective Hamiltonian
-$H_{\t{eff}}$ induced by $X_{\m w}$ on the ground-state manifold
-$\M_0$ through second order in perturbation theory
+eigenstates of $H_0$, which will in turn enable simulating the
+dynamics induced by non-perturbative additions $\hat\O$ to the SU($n$)
+Heisenberg model, with operator norms
+$\norm*{\hat\O}\sim\Delta_{\t{gap}}$.
+
+When the operator norm of $\hat\O$ is sufficiently small, namely
+$\norm*{\hat\O}<\Delta_{\t{gap}}/2$, we can treat its effect on the PS
+manifold $\M_0$ perturbatively.  The effective Hamiltonian
+$H_{\t{eff}}$ induced by $\hat\O$ on the ground-state manifold $\M_0$
+through second order in perturbation theory
 is\cite{bravyi2011schrieffer, perlin2019effective}
@@
   &&
-  H_{\t{eff}}^{(1)} = \P_0 X_{\m w} \P_0,
+  H_{\t{eff}}^{(1)} = \P_0 \hat\O \P_0,
   &&
   H_{\t{eff}}^{(2)} = - \sum_{\Delta\ne0}
-  \f1\Delta \P_0 X_{\m w} \P_\Delta X_{\m w} \P_0,
+  \f1\Delta \P_0 \hat\O \P_\Delta \hat\O \P_0,
   \label{eq:H_eff}
@@
 \begin{align}
-  H_{\t{eff}}^{(1)} \EQPS \bar w\,\col{X},
+  H_{\t{eff}}^{(1)} \EQPS \sum_{\p{X,\m w}\in\O} \bar w\,\col{X},
   \label{eq:H_eff_1}
@@
 where $\EQPS$ denotes equality under a restriction to $\M_0$; $\bar w$
-is an average of $w_k$ over spin choices $k\in\C_N\p{M}$; and
-$\col{X}$ is the collective version of $X$, i.e.
-\begin{align}
-  \bar w \equiv { N \choose M }^{-1} \sum_{k\in\C_N\p{M}} w_k,
-  &&
-  \col{X} \equiv \sum_{k\in\C_N\p{M}} X_k,
-\end{align}
-An immediate consequence of \eqref{eq:H_eff_1} is the fact that
-perturbing SU(2)-symmetric interactions by two-body $s_\z\otimes s_\z$
-interactions yields the one-axis twisting (OAT)
-Hamiltonian\cite{kitagawa1993squeezed, ma2011quantum}
-$H_{\t{OAT}}=\chi S_\z^2$ at first order in perturbation theory, with
-the OAT strength $\chi$ equal to the mean coefficient of the two-body
-$s_\z\otimes s_\z$ terms.
+is an average of $w_k$ over all $k\in\D_N\p{M}$; and $\col{X}$ is a
+permutation-symmetrized collective version of $X$, i.e.
+\begin{align}
+  \bar w \equiv \f1{\abs{\D_N\p{M}}} \sum_{k\in\D_N\p{M}} w_k,
+  &&
+  \col{X} \equiv \f1{\abs{\pi\p{M}}} \sum_{\sigma\in\pi\p{M}}
+  \sum_{k\in\C_N\p{M}} X_{\sigma\p{k}},
+\end{align}
+where, to be explicit,
+\begin{align}
+  \abs{\D_N\p{M}} = N!_M,
+  &&
+  \pi\p{M} = M!,
+\end{align}
+with $N!_M\equiv\prod_{j=0}^{M-1}\p{N-j}=M!\times{N\choose M}$
+denoting a falling factorial.  An immediate consequence of
+\eqref{eq:H_eff_1} is the fact that perturbing SU(2)-symmetric
+interactions by two-body $s_\z\otimes s_\z$ interactions yields the
+one-axis twisting (OAT) Hamiltonian\cite{kitagawa1993squeezed,
+  ma2011quantum} $H_{\t{OAT}}=\chi S_\z^2$ at first order in
+perturbation theory, with the OAT strength $\chi$ equal to the mean
+coefficient of the two-body $s_\z\otimes s_\z$ terms.
+
+\red{[TODO: fix remainder of this section to account for perturbations
+  that do not obey permutational symmetry]}
 
@@
 $\P_\Delta$ onto the manifold $\E_\Delta$ of states with excitation
-energy $\Delta$ (with respect to $H_0$).  This projector essentially
-picks off the part of $X_{\m w}$ that is strictly off-diagonal with
-respect to the ground- and excited-state manifolds $\E_0$ and
-$\E_\Delta$.  To decompose $X_{\m w}$ into components that generate
-states of definite excitation energy when acting on PS states $\M_0$,
-we pick an arbitrary state $\ket\psi\in\M_0$ and expand (see Appendix
-\ref{sec:eigenstates})
+energy $\Delta$ (with respect to $H_0$).  For each term $X_{\m w}$ of
+$\hat\O$, this projector essentially picks off the part of $X_{\m w}$
+that is strictly off-diagonal with respect to the ground- and
+excited-state manifolds $\E_0$ and $\E_\Delta$.  To decompose
+$X_{\m w}$ into components that generate states of definite excitation
+energy when acting on PS states $\M_0$, we pick an arbitrary state
+$\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:eigenstates})
 \begin{align}
@@
 $E_0+\Delta$.  It follows that
-$\P_\Delta X_{\m w} \P_0 = \P_\Delta X_{\m w_\Delta} \P_0 = X_{\m
-  w_\Delta} \P_0$, and so
-\begin{align}
-  H_{\t{eff}}^{(2)} = -\sum_{\Delta\ne0}
-  \f1\Delta \P_0 X_{\m w_\Delta} X_{\m w_\Delta} \P_0.
-\end{align}
-The product $\P_0 X_{\m w_\Delta} X_{\m w_\Delta} \P_0$ generally
+$\P_\Delta X_{\m w} \P_0 = \P_\Delta X_{\m w_\Delta} \P_0 =
+X_{\m{w}_\Delta} \P_0$, so
+\begin{align}
+  H_{\t{eff}}^{(2)} = -\sum_{\Delta\ne0} \f1\Delta
+  \sum_{\p{X,\m v},\p{Y,\m w}\in\O}
+  \P_0 X_{\m v_\Delta} Y_{\m w_\Delta} \P_0.
+\end{align}
+The product $\P_0 X_{\m v_\Delta} Y_{\m w_\Delta} \P_0$ generally
 includes $r$-body operators for all positive integers $r\le2M$.  We
@@
   H_{\t{eff}}^{(2)}
-  = \sum_{\Delta\ne0} \f{\var\p{\m w_\Delta}}{\p{N-1}\Delta}
-  \p{\col{X}^2 - N \col{X^2}},
-  &&
-  \var\p{\m m} \equiv \f1N \sum_{p\in\ZZ_N} \p{m_p-\bar m}^2,
-\end{align}
-which in the SU(2) case implies that an inhomogeneous magnetic field
-($X\propto\sigma_\z$) yields an OAT Hamiltonian at second order
+  = \sum_{\Delta\ne0} \sum_{\p{X,\m v},\p{Y,\m w}\in\O}
+  \f{\cov\p{\m v_\Delta,\m w_\Delta}}{\p{N-1}\Delta}
+  \p{\col{X}\,\col{Y} - N \col{XY}},
+  \label{eq:H_eff_2_1}
+\end{align}
+where
+\begin{align}
+  \cov\p{\m v,\m w} \equiv
+  \f1N \sum_{p\in\ZZ_N} \p{v_p-\bar v}\p{w_p-\bar w}
+\end{align}
+is the covariance between $\m v$ and $\m w$.  In the SU(2) case,
+\eqref{eq:H_eff_2_1} implies that an inhomogeneous magnetic field
+($X\propto s_\z$) yields an OAT Hamiltonian at second order
 perturbation theory.
@@
 In order to find the effective Hamiltonian $H_{\t{eff}}^{(2)}$ induced
-by the $M$-body operator $X_{\m w}$ on the PS manifold $\M_0$ at
+by the multi-body operator $\hat\O$ on the PS manifold $\M_0$ at
 second order in perturbation theory, we constructed states with
@@
 
+\red{[TODO: fix section to consider the case without permutational
+  symmetry of $X_{\m w}$]}
+
 In this section, we simplify the products of the form
@@
 \begin{align}
-  X_{\m w} \equiv \sum_{k\in\C_N\p{M}} w_k X_k,
+  X_{\m w} \equiv \f1{M!} \sum_{k\in\D_N\p{M}} w_k X_k,
   \label{eq:pert_copy}
@@
 where $X$ is an $M$-spin operator; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is
-an ordered choice of $M$ spins that $X_k$ acts on; $\C_N\p{M}$ is the
-set of all such choices of $M$ elements from a set of size $N$; and
-$\m w$ is a dimension-$M$ (i.e.~$M$-index) tensor with scalar entries
-$w_k$.  Even though only ${N\choose M}$ entries of $\m w$ are used to
-define $X_{\m w}$ in \eqref{eq:pert_copy}, for later convenience we
-define $w_k$ for all indices $k\in\ZZ_N^M$ by enforcing that $\m w$ is
-symmetric with respect to any permutation of its indices, and that
-$w_k=0$ for any $k$ with repeated values.  In contrast, $X_k$
-generally need not be symmetric with respect to permutations of $k$,
-although we will still define $X_k=0$ for any $k$ with repeated
-values.
+a list of $M$ spins that $X_k$ acts on; $\D_N\p{M}$ is defined in
+\eqref{eq:off_diags}; and $\m w$ is a dimension-$M$ (i.e.~$M$-index)
+tensor with scalar entries $w_k$.
 
@@
 
-\red{[TODO: verify that the derivation below properly accounts for
-  operators $X$ that are not permutationally symmetric]}
+\red{[TODO: fix section to consider the case without permutational
+  symmetry of $X_{\m w}$]}
 
@@
 permutation operator $\Pi_{pq}$ acts trivially on $X_k\ket\psi$ when
-$p,q\notin k$.  When $p,q\in k$, this permutation operator swaps the
-positions of $p$ and $q$ in $X_k$, but we can simply relabel
-$\p{p,q}\leftrightarrow\p{q,p}$ with no consequence for $w_k$, so
+$p,q\in k$ or $p,q\notin k$, so
 \begin{align}
   \sum_{\substack{k\in\C_N\p{M}\\\p{p,q}\in\C_N\p{2}\\
-      p,q\notin k~\t{or}~p,q\in k}}
+      p,q\in k~\t{or}~p,q\notin k}}
   h_{pq} w_k \Pi_{pq} X_k \ket\psi
   = \sum_{k\in\C_N\p{M}}
-  \sum_{\substack{\p{p,q}\in\C_N\p{2}\\p,q\notin k~\t{or}~p,q\in k}}
+  \sum_{\substack{\p{p,q}\in\C_N\p{2}\\p,q\in k~\t{or}~p,q\notin k}}
   h_{pq} w_k X_k \ket\psi.
 \end{align}
-For fixed $k\in\C_N\p{M}$, we can simplify
+For any fixed $k\in\C_N\p{M}$, we can simplify
 \begin{align}
@@
 Given a system of $N$ spins, we wish to project a product of
-multi-body operators onto the permutationally symmetric manifold (PS)
+multi-body operators onto the permutationally symmetric (PS) manifold
 $\M_0$.  For $p$ multi-body operators, such a product takes the form
@@
   \m X_{\m w}
-  \equiv \prod_{\alpha\in\ZZ_p} \sum_{k_\alpha\in\C_N\p{M_\alpha}}
+  \equiv \prod_{\alpha\in\ZZ_p} \f1{M_\alpha!}
+  \sum_{k_\alpha\in\D_N\p{M_\alpha}}
   w_\alpha\p{k_\alpha} X_\alpha\p{k_\alpha},
@@
 $k_\alpha\equiv\p{k_{\alpha,1},k_{\alpha,2},\cdots,k_{\alpha,M_\alpha}}$
-is an ordered choice of $M_\alpha$ spins that $X_\alpha\p{k_\alpha}$
-acts on; $\C_N\p{M_\alpha}$ is the set of all such choices of
-$M_\alpha$ elements from a set of size $N$; and $w_\alpha\p{k_\alpha}$
-is a scalar entry of $w_\alpha$.  Although only ${N \choose M_\alpha}$
-entries of $w_\alpha$ are used in \eqref{eq:sym_prod_start}, we define
-$w_\alpha\p{k_\alpha}$ for all $k_\alpha\in\ZZ_N^{M_\alpha}$ by
-enforcing that $w_\alpha\p{k_\alpha}$ is symmetric with respect to
-permutations of $k_\alpha$, and that $w_\alpha\p{k_\alpha}=0$ if
-$k_\alpha$ has any repeated values.  In contrast,
-$X_\alpha\p{k_\alpha}$ generally need not be symmetric with respect to
-permutations of $k_\alpha$, although we still enforce that
-$X_\alpha\p{k_\alpha}=0$ if $k_\alpha$ has any repeated values.
-
-To simplify the product in \eqref{eq:sym_prod_start}, we rewrite the
-product of sums as a sum of products, and symmetrize over all
-permutations of $k_\alpha\in\C_N\p{M_\alpha}$ in order to make all
-individual sums independent:
+is a choice of $M_\alpha$ spins that $X_\alpha\p{k_\alpha}$ acts on;
+$\D_N\p{M_\alpha}$ is the set of all ordered choices of $M_\alpha$
+elements from a set of size $N$; and $w_\alpha\p{k_\alpha}$ is a
+scalar entry of $w_\alpha$.  For convenience, we define
+$w_\alpha\p{k_\alpha}$ and $X_\alpha\p{k_\alpha}$ for all
+$k_\alpha\in\ZZ_N^{M_\alpha}$ by enforcing that
+$w_\alpha\p{k_\alpha}=X_\alpha\p{k_\alpha}=0$ for all $k_\alpha$ with
+repeated values.  These definitions allow us to replace sums over
+$\D_N\p{M_\alpha}$ by sums over $\ZZ_N^{M_\alpha}$, thereby making the
+sums over the individual indices in $k_\alpha$ independent.
+
+To simplify the product in \eqref{eq:sym_prod_start}, we first replace
+all $\D_N\p{M_\alpha}$ by $\ZZ_N^{M_\alpha}$ and rewrite the product
+of sums as a sum of products:
 \begin{align}
@@
   &= \f1{M_1!} \sum_{k_1\in\ZZ_N^{M_1}}
-  \f1{M_2!} \sum_{k_2\in\ZZ_N^{M_2}} \cdots
+  \f1{M_2!} \sum_{k_2\in\ZZ_N^{M_1}} \cdots
   \f1{M_p!} \sum_{k_p\in\ZZ_N^{M_p}}
@@
 \end{align}
-where $\m M\equiv\p{M_1,M_2,\cdots,M_p}$ is a vector of the
-dimensions $M_\alpha$; $\m M!\equiv\prod_{M_\alpha\in\m M}M_\alpha!$
-for shorthand; and
+where $\m M\equiv\p{M_1,M_2,\cdots,M_p}$ is a vector of the dimensions
+$M_\alpha$; $\m M!\equiv\prod_{M_\alpha\in\m M}M_\alpha!$ for
+shorthand; and
 $\ZZ_N^{\m M}\equiv\bigotimes_{\alpha\in\ZZ_p}\ZZ_N^{M_\alpha}$, such
@@
 diagram $g$), and we denote the set of all diagrams for
-$p\equiv\dim\p{\m M}$ tensors with dimensions $\m M$ by $\G\p{\m M}$.
-For consistency, all diagrams $g\in\G\p{\m M}$ must associate exactly
+$p=\dim\p{\m M}$ tensors with dimensions $\m M$ by $\G\p{\m M}$.  For
+consistency, all diagrams $g\in\G\p{\m M}$ must associate exactly
 $M_\alpha$ indices with the dimension-$M_\alpha$ tensor $w_\alpha$,
@@
 \end{align}
-for all $\alpha\in\ZZ_p$.  Schematically, a classification of the
-terms in \eqref{eq:sum_of_products} by diagrams $g\in\G\p{\m M}$ will
-allow us to expand
-\begin{align}
-  \m X_{\m w} \EQPS \sum_{g\in\G\p{\m M}} \m w\p{g} \m X\p{g},
-  \label{eq:sym_prod_teaser}
+for all $\alpha\in\ZZ_p$.
+
+Schematically, a classification of the terms in
+\eqref{eq:sum_of_products} by diagrams $g\in\G\p{\m M}$ will allow us
+to expand
+\begin{align}
+  \m X_{\m w} \EQPS \sum_{g\in\G\p{\m M}} \m X_{\m w}\p{g},
+  \label{eq:sym_prod_diagrams}
 \end{align}
 where $\EQPS$ denotes equality up to a restriction to the PS manifold
-$\M_0$; $\m w\p{g}$ is a scalar acquired by assigning $\abs{g}$
-indices to $\m w$ according to $g$, and summing over those indices in
-some appropriate manner; and $\m X\p{g}$ is a $\abs{g}$-spin operator
-acquired by averaging over all of the distinct ways to assign
-$\abs{g}$ indices to $\m X$ according to $g$ (clarified below).
-
-In order to derive the expansion in \eqref{eq:sym_prod_teaser} and
-find the explicit forms of $\m w\p{g}$ and $\m X\p{g}$, we need to
-collect all terms in \eqref{eq:sum_of_products} that are associated
-with the diagram $g$.  To this end, we first consider all distinct
-ways to assign $\abs{g}$ indices to specific tensor factors of $\m w$
-and $\m X$ in a manner consistent with $g$.  For example, if
-$\m X=\p{X_1,X_2}$ with dimensions $\p{M_1,M_2}=\p{2,1}$ and
-$\abs{g}=2$ with $g_{\set{1,2}}=g_{\set{2}}=1$, then
-$X_1\p{a,b} X_2\p{b}$ and $X_1\p{b,a} X_2\p{b}$ are distinct
-assignments of $\abs{g}$ indices to $\m X$ in a manner consistent with
-$g$, provided that $a\ne b$.  After assigning $\abs{g}$ indices to
-$\m w$ and $\m X$ in a manner consistent with $g$, one needs to sum
-over all values that these indices can take, which amounts to a sum
-over all $k=\p{k_1,k_2,\cdots,k_{\abs{g}}}\in\ZZ_N^{\abs{g}}$ with the
-restriction that $k_\mu\ne k_\nu$ for all $\mu\ne\nu$.
+$\M_0$; and $\m X_{\m w}\p{g}$ is the sum of all terms in
+\eqref{eq:sum_of_products} that are associated with the diagram $g$.
+To collect these terms, we first consider all distinct ways to assign
+$\abs{g}$ indices to specific tensor factors of $\m w$ and $\m X$ in a
+manner consistent with $g$.  For example, if $\m X=\p{X_1,X_2}$ with
+dimensions $\p{M_1,M_2}=\p{2,1}$ and $\abs{g}=2$ with
+$g_{\set{1,2}}=g_{\set{2}}=1$, then $X_1\p{a,b} X_2\p{b}$ and
+$X_1\p{b,a} X_2\p{b}$ are distinct assignments of $\abs{g}$ indices to
+$\m X$ in a manner consistent with $g$, provided that $a\ne b$.  After
+assigning $\abs{g}$ indices to $\m w$ and $\m X$ in a manner
+consistent with $g$, we need to sum over all values that these indices
+can take, which amounts to a sum over all
+$k=\p{k_1,k_2,\cdots,k_{\abs{g}}}\in\ZZ_N^{\abs{g}}$ with the
+restriction that $k_\mu\ne k_\nu$ for all $\mu\ne\nu$ to enforce that
+all indices are distinct, or equivalently a sum over all
+$k\in\D_N\p{\abs{g}}$.
 
@@
 The assignment $X_1\p{a,b,c} X_2\p{a,b}$, for example, is the same as
-$X_1\p{b,a,c} X_2\p{b,a}$.  In general, the treatment of the $g_S$
-indistinguishable dots within region $S\in\PP^*\p{\ZZ_p}$ of $g$ as
-distinguishable results in over-counting distinct index assignments by
-a factor of $g_S!$.  Denoting the set of all distinct index
-assignments according to $g$ by $\L\p{g}$, the number of such
-assignments is therefore
-\begin{align}
-  \abs{\L\p{g}} = \m M! \sp{\prod_{S\in\PP^*\p{\ZZ_p}} g_S!}^{-1}.
-\end{align}
+$X_1\p{b,a,c} X_2\p{b,a}$, because $a$ and $b$ are dummy indices.  In
+general, the treatment of the $g_S$ indistinguishable dots within
+region $S\in\PP^*\p{\ZZ_p}$ of $g$ as distinguishable results in
+over-counting distinct index assignments by a factor of $g_S!$.
+Denoting the set of all distinct index assignments according to $g$ by
+$\L\p{g}$, the number of such assignments is therefore
+\begin{align}
+  \abs{\L\p{g}} = \f{\m M!}{\S\p{g}},
+  &&
+  \S\p{g} \equiv \prod_{S\in\PP^*\p{\ZZ_p}} g_S!,
+\end{align}
+where $\S\p{g}$ is a symmetry factor associated with the diagram $g$.
 Altogether, we have
 \begin{align}
-  \m X_{\m w}
-  = \f1{\m M!} \sum_{g\in\G\p{\m M}} \sum_{\ell\in\L\p{g}}
-  \sum_{\substack{k\in\ZZ_N^{\abs{g}}\\
-      k_\mu\ne k_\nu\forall \mu\ne\nu}}
-  \m w\p{k,\ell} \m X\p{k,\ell},
+  \m X_{\m w}\p{g} \equiv \f1{\m M!} \sum_{\ell\in\L\p{g}}
+  \sum_{k\in\D_N\p{\abs{g}}} \m w\p{k,\ell} \m X\p{k,\ell},
 \end{align}
@@
 indices $k$ to all $w_\alpha$ and $X_\alpha$ according to $\ell$.  The
-invariance of each tensor $w_\alpha$ under arbitrary permutations of
-its indices means that all assignments of $\abs{g}$ indices to $\m w$
-consistent with $g$ are equivalent,
-e.g.~$w_1\p{a,b} w_2\p{b} = w_1\p{b,a} w_2\p{b}$.  The value of
-$\m w\p{k,\ell}$ therefore depends on $g$, but not the specific choice
-of $\ell\in\L\p{g}$.  Furthermore, the specific choice of spins $k$ in
-$\m X\p{k,\ell}$ does not matter after projection onto the PS manifold
-$\M_0$, so the dependence of $\m X\p{k,\ell}$ on $k$ is similarly
-superfluous, which is to say that we can write
-\begin{align}
-  \m X_{\m w}
-  \EQPS \f1{\m M!} \sum_{g\in\G\p{\m M}}
-  \sum_{\substack{k\in\ZZ_N^{\abs{g}}\\
-      k_\mu\ne k_\nu\forall \mu\ne\nu}}
-  \m w\p{k,g} \sum_{\ell\in\L\p{g}} \m X\p{\ell}.
-\end{align}
-Defining the average of $\m X\p{\ell}$ over all $\ell\in\L\p{g}$,
-\begin{align}
-  \m X\p{g} \equiv \f1{\abs{\L\p{g}}}
-  \sum_{\ell\in\L\p{g}} \m X\p{\ell},
-\end{align}
-we can more simply write
-\begin{align}
-  \m X_{\m w}
-  \EQPS \f1{\m M!} \sum_{g\in\G\p{\m M}} \abs{\L\p{g}}
-  \sum_{\substack{k\in\ZZ_N^{\abs{g}}\\
-      k_\mu\ne k_\nu\forall \mu\ne\nu}}
-  \m w\p{k,g} \m X\p{g}.
-\end{align}
-We now examine the sum over all $k\in\ZZ_N^{\abs{g}}$ with
-$k_\mu\ne k_\nu\forall \mu\ne\nu$.  For any given instance of
-$\m w\p{k,g}$, each entry $k_\mu\in k$ is associated with tensors
-$\set{w_\alpha:\alpha\in S}$ for some subset $S\in\PP^*\p{\ZZ_p}$.
-Identifying the indices associated with $S\in\PP^*\p{\ZZ_p}$ by
-$k_S\in\ZZ_N^{g_S}$ and enforcing the decomposition
-$k=\p{k_S:S\in\PP^*\p{\ZZ_p}}$, we can write
-\begin{align}
-  \sum_{\substack{k\in\ZZ_N^{\abs{g}}\\
-      k_\mu\ne k_\nu\forall \mu\ne\nu}} \m w\p{k,g}
-  = \sum_{k\in\ZZ_N\p{g}} \m w\p{k},
-\end{align}
-where
-\begin{align}
-  \ZZ_N\p{g} \equiv
-  \set{ k \in \bigotimes_{S\in\PP^*\p{\ZZ_p}} \ZZ_N^{g_S}
-    : \t{all elements of $k$ are distinct} }.
-\end{align}
-The domain $\ZZ_N\p{g}$ is redundant, however, in the sense that
-$\m w\p{k}$ for any $k\in\ZZ_N\p{g}$ is invariant under permutations
-of each $k_S\in\ZZ_N\p{g_S}$, so
-\begin{align}
-  \sum_{k\in\ZZ_N\p{g}} \m w\p{k}
-  = \sp{\prod_{S\in\PP^*\p{\ZZ_p}}g_S!} \sum_{k\in\C_N\p{g}} \m w\p{k},
+specific choice of spins $k$ in $\m X\p{k,\ell}$ does not matter after
+projection onto the PS manifold $\M_0$, so the dependence of
+$\m X\p{k,\ell}$ on $k$ is superfluous, which is to say that we can
+write
+\begin{align}
+  \m X_{\m w}\p{g} \EQPS \f1{\m M!} \sum_{\ell\in\L\p{g}}
+  \sum_{k\in\D_N\p{\abs{g}}} \m w\p{k,\ell} \m X\p{\ell},
+\end{align}
+where $\EQPS$ denotes equality under a restriction to $\M_0$.  We now
+define, for any index assignment $\ell\in\L\p{g}$, the scalar
+\begin{align}
+  \m w\p{\ell}
+  \equiv \f1{\S\p{g}} \sum_{k\in\D_N\p{\abs{g}}} \m w\p{k,\ell},
+  \label{eq:assignment_scalar}
+\end{align}
+where the factor $1/\S\p{g}$ is included to average over the distinct
+ways to permute the shared indices in $\set{w_\alpha:\alpha\in S}$ for
+all $S\in\PP^*\p{\ZZ_p}$, which is essentially an average over the
+symmetries of $g$.  The motivation for this averaging comes from the
+consideration of the special case in which all $w_\alpha$ are
+invariant under arbitrary permutations of their tensor factors (or
+simply ``permutationally symmetric''),
+i.e.~$w_\alpha\p{k}=w_\alpha\p{\sigma\p{k}}$ for all permutations
+$\sigma\in\pi\p{M_\alpha}$.  In this case, we can simplify
+\begin{align}
+  \m w\p{\ell} = \sum_{k\in\C_N\p{g}} \m w\p{k,\ell},
 \end{align}
@@
   \set{ k \in \bigotimes_{S\in\PP^*\p{\ZZ_p}} \C_N\p{g_S}
-    : \t{all elements of $k$ are distinct} }.
-\end{align}
-Defining
-\begin{align}
-  \m w\p{g} \equiv \sum_{k\in\C_N\p{g}} \m w\p{k},
-  \label{eq:diagram_factors}
-\end{align}
-we thus find that
-\begin{align}
-  \m X_{\m w} \EQPS \sum_{g\in\G\p{\m M}} \m w\p{g} \m X\p{g},
-  \label{eq:sym_prod_diagrams}
-\end{align}
-where all combinatorial factors such as $\m M!$ and $\abs{\L\p{g}}$
-cancel each other out.
+    : \t{all elements of $k$ are distinct} }
+\end{align}
+is essentially the set of all index values $j,k\in\C_N\p{g}$ for which
+generally $\m w\p{j,\ell}\ne\m w\p{k,\ell}$ if $j\ne k$.  Note that if
+all $w_\alpha$ are permutationally symmetric, then $\m w\p{\ell}$
+depends on the diagram $g$, but not on the specific choice of index
+assignment $\ell\in\L\p{g}$.
+
+The definition of $\m w\p{\ell}$ in \eqref{eq:assignment_scalar}
+allows us to expand
+\begin{align}
+  \m X_{\m w}\p{g} \EQPS \f1{\abs{\L\p{g}}}
+  \sum_{\ell\in\L\p{g}} \m w\p{\ell} \m X\p{\ell},
+  \label{eq:diagram_terms}
+\end{align}
+where if additionally all tensors in $\m w$ {\it or} all operators in
+$\m X$ are permutationally symmetric, then
+\begin{align}
+  \m X_{\m w}\p{g} \EQPS \m w\p{g} \m X\p{g},
+  \label{eq:diagram_terms_perm}
+\end{align}
+with
+\begin{align}
+  \m w\p{g} \equiv \f1{\abs{\L\p{g}}}
+  \sum_{\ell\in\L\p{g}} \m w\p{\ell}
+  &&
+  \m X\p{g} \equiv \f1{\abs{\L\p{g}}}
+  \sum_{\ell\in\L\p{g}} \m X\p{\ell}.
+  \label{eq:assignment_averages}
+\end{align}
+The simplification in \eqref{eq:diagram_terms_perm} also holds if all
+$X_\alpha$ are permutationally symmetric, in which case $\m w\p{g}$ is
+similarly defined as an average of $\m w\p{\ell}$ over all
+$\ell\in\L\p{g}$.
+
+In summary, the product of operators $\m X_{\m w}$ in
+\eqref{eq:sym_prod_start} admits the diagrammatic expansion in
+\eqref{eq:sym_prod_diagrams}, as a sum of $\m X_{\m w}\p{g}$ over all
+diagrams $g\in\G\p{\m M}$.  Each term $\m X_{\m w}\p{g}$ is
+essentially an average of $\m w\p{\ell} \m X\p{\ell}$ over all
+distinct index assignments $\ell\in\L\p{g}$, as in
+\eqref{eq:diagram_terms}.  The scalar $\m w\p{\ell}$ is obtained by
+assigning indices $k=\p{k_1,k_2,\cdots,k_{\abs{g}}}$ to the tensors
+$\m w$ as prescribed by the $\ell$, and summing over all
+$k\in\D_N\p{\abs{g}}$, as in \eqref{eq:assignment_scalar}.  The
+$\abs{g}$-spin operator $\m X\p{\ell}$ is obtained by contracting the
+$M_\alpha$-spin operators $X_\alpha$ in $\m X$ in accordance with the
+index assignment $\ell$.  If the tensors $\m w$ or operators $\m X$
+are permutationally symmetric, then $\m X_{\m w}\p{g}$ factorizes into
+$\m w\p{g} \m X\p{g}$, where $\m w\p{g}$ and $\m X\p{g}$ are
+respectively defined as the average of $\m w\p{\ell}$ and
+$\m X\p{\ell}$ over all $\ell\in\L\p{g}$.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Formalizing diagrams and reducing computational costs}
+\subsection{Formalizing diagrams}
 \label{sec:diagrams}
 
-Diagrams such as \eqref{eq:venn_diagram} will play a major role in the
-expansion in \eqref{eq:sym_prod_diagrams}, as well as the remainder of
-this appendix.  Here, we define these diagrams as formal objects that
-we can use in mathematical expressions, enabling us to systematically
-perform calculations that are otherwise intractable.  We then derive
-rules to ``simplify'' these diagrams in such a way as to reduce the
-computational cost of their numerical evaluation.
-
-The simplest diagrams have $g_S$ filled dots (``$\bullet$'') at the
-intersection of the circles $j\in S$, and are simply equal to the
-associated scalar $\m w\p{g}$ defined in \eqref{eq:diagram_factors},
-e.g.
+Diagrams such as \eqref{eq:venn_diagram} play a central role in
+simplifying products of multi-body operators, through the expansion in
+\eqref{eq:sym_prod_diagrams}.  Here, we define these diagrams as
+formal objects that enable us to systematically perform otherwise
+intractable calculations.  We then derive rules to ``reduce'' these
+diagrams in such a way as to minimize the computational cost of their
+numerical evaluation.
+
+All diagrams consist of $p$ circles, and label each circle
+$\alpha\in\ZZ_p$ by some tensor $w_\alpha$, as in
+\eqref{eq:venn_diagram} (or below).  The simplest diagrams have $g_S$
+filled dots (``$\bullet$'') at the intersection of the circles
+$\alpha\in S\subset\ZZ_p$, and are simply equal to the scalar
+$\m w\p{g}$ defined in \eqref{eq:assignment_averages}.  If all tensors
+$w_\alpha$ are permutationally symmetric, for example, then
 \begin{align}
   \diagram{example_123}
-  \equiv \sum_{\substack{a,b,c,\in\ZZ_N\\\p{d,e}\in\C_N\p{2}\\
+  = \sum_{\substack{a,b,c,\in\ZZ_N\\\p{d,e}\in\C_N\p{2}\\
       \abs{\set{a,b,c,d,e}}=5}}
   w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e},
+  \label{eq:example_123}
 \end{align}
 where the restriction $\abs{\set{a,b,c,d,e}}=5$ denotes that all of
-$a,b,c,d,e$ are distinct.  Such a diagram with $\abs{g}$ dots
-nominally takes $O\p{N^{\abs{g}}}$ time to compute.  In practice, this
-computational cost can be prohibitive for performing numerical
-simulations of systems with any appreciable size.  We therefore wish
-to simplify diagrams in order to reduce their computational cost as
-much as possible.  Our general strategy will be to replace constrained
-sums that make all indices interdependent by unconstrained sums that
-can be carried out independently, in sequence.  In order to represent
-different constraints, we need to define diagrams in which filled dots
-(``$\bullet$'') may be replaced by empty dots (``$\circ$'') or crosses
-(``$\bm\times$'').
+$a,b,c,d,e$ are distinct.  When the tensors $w_\alpha$ are not all
+permutationally symmetric, one should generally equip each diagram
+with an index assignment $\ell\in\L\p{g}$, in which case the value of
+this diagram is equal to the scalar $\m w\p{\ell}$ defined in
+\eqref{eq:assignment_scalar}; if no such assignment is provided, then
+the diagram is equal to the average over all index assignments
+consistent with $g$.  To keep our discussions simple, we will continue
+to assume that all tensors $w_\alpha$ are permutationally symmetric,
+but note that all discussions generally apply to the asymmetric case
+as well, as long as all sums over $\C_N\p{M}$ are replaced by a sums
+over $\D_N\p{M}$ with a symmetry prefactor of $1/M!$,
+i.e.~$\sum_{k\in\C_N\p{M}}f\p{k} \to
+\p{1/M!}\sum_{k\in\D_N\p{M}}f\p{k}$.  The only exception, in which
+diagrams equipped with index assignments $\ell\in\L\p{g}$ are treated
+differently than diagrams without such assignments, concerns one step
+in the process of ``reducing'' diagrams (specifically, the step of
+``eliminating crosses'').  We will comment on this exception when it
+appears.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Reducing computational costs}
+
+A diagram such as \eqref{eq:example_123} with $\abs{g}$ dots nominally
+takes $O\p{N^{\abs{g}}}$ time to compute due to the interdependence of
+all indices.  We wish to ``reduce'' such diagrams and minimize their
+computational cost as much as possible.  Our general strategy will be
+to replace constrained sums that make all indices interdependent by
+unconstrained sums that can be carried out independently, in sequence.
+In order to represent different constraints, we need to define
+diagrams in which filled dots (``$\bullet$'') may be replaced by empty
+dots (``$\circ$'') or crosses (``$\bm\times$'').
 
@@
   \diagram{example_o}
-  \equiv \sum_{\substack{\p{a,b}\in\C_N\p{2}\\e\in\ZZ_N\\
+  = \sum_{\substack{\p{a,b}\in\C_N\p{2}\\e\in\ZZ_N\\
       \abs{\set{a,b,e}}=3}}
@@
   \diagram{example_x}
-  \equiv \sum_{\substack{a,e\in\ZZ_N\\\abs{\set{a,e}}=2}}
+  = \sum_{\substack{a,e\in\ZZ_N\\\abs{\set{a,e}}=2}}
   \sum_{b,c\in\set{a,e}} \sum_{d\in\ZZ_N}
@@
 computational cost into different diagrams with a lower computational
-cost.  We will essentially use two tricks to decompose diagrams:
+cost.  We will essentially use two tricks to reduce diagrams:
 eliminating filled dots (which will add empty dots and crosses), and
-eliminating crosses (which will move around filled dots).
+eliminating crosses (which will move around filled dots).  These two
+tricks will be applied in sequence until only empty dots remain.
 
@@
   = \f14 \diagram{example_sym_o}
-  - \f14 \diagram{example_sym_x}.
-\end{align}
-This factor can be derived by replacing the sum over $\C_N\p{F}$ for
-the filled dots within the region by a sum over $\ZZ_N^{F}$, which
-picks up a factor of $1/F!$ to account for the number of ways to
-permute $F$ values.  Setting aside one filled dot for replacement by
-an empty dot and a cross, the sum over $\ZZ_N^{F-1}$ for the remaining
-dots can be changed back to a sum over $\C_N\p{F-1}$, picking up a
-factor of $\p{F-1}!$.  The overall factor picked up throughout this
-procedure is therefore $\p{F-1}!/F!=1/F$.  Similarly, adding an empty
-dot to a region with $E$ empty dots makes the diagram pick up a factor
-of $\p{E+1}!/E!=E+1$, so
+  - \f14 \diagram{example_sym_x},
+\end{align}
+This factor simply accounts for the fact that the diagram on the left
+has an overall symmetry factor of $1/4!\times1/4!$, while the diagrams
+on the right only have an overall symmetry factor of $1/3!\times1/4!$.
+Similarly, adding an empty dot to a region with $E$ empty dots makes
+the diagram pick up a factor of $E+1$, so
 \begin{align}
@@
 where the factor of $1/F_{\t{old}}$ that is acquired due to removing a
-dot from the old region is canceled out by a factor of $F_{\t{old}}$
-that accounts for the number of ways to move a dot from the old region
-to the new one.
+dot from the old region is, loosely speaking, canceled out by the
+$F_{\t{old}}$ choices of which dot to move.  More precisely,
+eliminating a cross and moving a dot from the old region to the new
+one corresponds to enforcing that the index associated with the cross
+take on the same value as an index associated with a filled dot in the
+old region, and there are $F_{\t{old}}$ such indices to choose from.
+
+If the tensors $\set{w_\alpha:\alpha\in\ZZ_p}$ associated with a
+diagram are not symmetric, then all dots in a given region
+$S\subset\ZZ_p$ of a diagram are distinguishable by their association
+with a specific tensor factor of each $w_\alpha$ with $\alpha\in S$.
+Rather than picking up a factor of $F_{\t{old}}$ to account for the
+number of ways to move a dot from the old region to the new one, one
+should then sum over each choice of which dot to move, e.g.
+\begin{align}
+  \diagram{example_sym_x_color}
+  = \f13 \sp{\diagram{example_sym_x_color_r}
+    + \diagram{example_sym_x_color_m}
+    + \diagram{example_sym_x_color_b}}.
+\end{align}
 
@@
 
+\red{[TODO: fix section to consider the case without permutational
+  symmetry]}
+
 Here we simplify the two-body product
@@
 
+\red{[TODO: fix section to consider the case without permutational
+  symmetry]}
+
 Here we simplify the product

### 2020-03-24 collective_sun/scripts/make_diagrams.py

@@ -70,9 +70,16 @@ region_mids["ac"] = rotate("b") @ region_mids["bc"]
 ##################################################
 # define constructor for diagram
 
-def place_dots(region, num, markers = None, dot_sep = None,
+def place_dots(region, num_colors, markers = None, dot_sep = None,
                region_mids = region_mids, markersize = 3,
                default_marker = None):
+    if type(num_colors) is int:
+        num = num_colors
+        colors = [ "k" for _ in range(num) ]
+    else:
+        num, colors = num_colors
+        assert(len(colors) >= num)
+
     if default_marker is None:
         default_marker = "."
     if dot_sep == None:
@@ -107,7 +114,7 @@ def place_dots(region, num, markers = None, dot_sep = None,
         exit()
 
     point = region_mids[region]
-    for vec, marker in zip(vecs, markers):
+    for vec, marker, color in zip(vecs, markers, colors):
         size = markersize
         markerfacecolor = "k"
 
@@ -122,7 +129,7 @@ def place_dots(region, num, markers = None, dot_sep = None,
             size += 1
 
         pos = np.array(point) + np.array(vec)
-        plt.plot(*pos, "k"+marker, markersize = size, fillstyle = fill_style)
+        plt.plot(*pos, marker, color = color, markersize = size, fillstyle = fill_style)
 
 def make_triple_diagram(region_dots, markers = {}, figsize = figsize,
                         dot_sep = None, pad = radius/20,
@@ -198,13 +205,15 @@ def make_double_diagram(region_dots, markers = {}, size_x = figsize[1],
     plt.tight_layout(pad = 0)
     return fig
 
-def add_triple_labels(labels, pads = [ 0.3, 0.4, 0.4 ]):
+def add_triple_labels(labels = None, pads = [ 0.3, 0.4, 0.4 ]):
+    if labels is None: labels = [ "u", "v", "w" ]
[…]

### 2020-03-25 collective_sun/sun_perturbation.tex
finalized conventions with symmetry factors, etc. in the definition and product of multi-body operators

@@
 \newcommand{\T}{\mathcal{T}}
-\newcommand{\Z}{\mathcal{Z}}
 
@@
 \begin{align}
-  H_0 = \sum_{\p{p,q}\in\C_N\p{2}} h_{pq} \Pi_{pq},
+  H_0 = \sum_{\substack{p,q\in\ZZ_N\\p<q}} h_{pq} \Pi_{pq},
   &&
@@
 \end{align}
-where $p,q$ index individual spins; $\C_N\p{M}$ is the set of all
-choices of $M$ spins out of $N$; $h_{pq}$ are scalar coupling
-constants; $\mu,\nu$ index states in an orthonormal basis for the
-$n$-dimensional Hilbert space of a single spin; $\ZZ_N$ is the set of
-integers modulo $N$; the operator
-$s_{\mu\nu}^{(p)}\equiv\op{\mu}{\nu}_p$ flips the state of spin $p$ to
-$\ket\mu$ from $\ket\nu$; and the operator $\Pi_{pq}$ permutes the
-states of spins $p$ and $q$.  The permutation operator $\Pi_{pq}$ is
-essentially a multilevel generalization of the SU(2)-symmetric spin
-interaction $\v s_p\c\v s_q$, with $\v s\equiv\p{\v s_\x,s_\y,s_\z}$,
+where $p,q$ index individual spins; $\ZZ_k$ is the set of integers
+modulo $k$; $h_{pq}$ are scalar coupling constants; $\mu,\nu$ index
+states in an orthonormal basis for the $n$-dimensional Hilbert space
+of a single spin; the operator $s_{\mu\nu}^{(p)}\equiv\op{\mu}{\nu}_p$
+flips the state of spin $p$ to $\ket\nu$ from $\ket\mu$; and the
+operator $\Pi_{pq}$ permutes the states of spins $p$ and $q$.  The
+permutation operator $\Pi_{pq}$ is essentially a multilevel
+generalization of the SU(2)-symmetric spin interaction
+$\v s_p\c\v s_q$, with $\v s\equiv\p{\v s_\x,s_\y,s_\z}$,
 $s_\alpha\equiv\sigma_\alpha/2$, and $\sigma_\alpha$ a single-spin
@@
 interaction energy $\Delta_{\t{gap}}>0$.  Power-law couplings of the
-form $h_{pq}\sim-1/\abs{p-q}^\alpha$ on a $D$-dimensional lattice, for
-example, always yield a non-vanishing spectral gap $\Delta_{\t{gap}}$
-for long-range interactions with $\alpha\le D$ (see Appendix
-\ref{sec:spectral_gap}).
+form $h_{pq}\sim-1/\abs{r_p-r_q}^\alpha$ with positions $r_p,r_q$ on a
+$D$-dimensional lattice, for example, always yield a non-vanishing
+spectral gap $\Delta_{\t{gap}}$ for long-range interactions with
+$\alpha\le D$ (see Appendix \ref{sec:spectral_gap}).
 
@@
 \begin{align}
-  E_0 = \sum_{\p{p,q}\in\C_N\p{2}} h_{pq}.
+  E_0 = \sum_{\substack{p,q\in\ZZ_N\\p<q}} h_{pq}.
 \end{align}
@@
   &&
-  X_{\m w} \equiv \f1{M!} \sum_{k\in\D_N\p{M}} w_k X_k,
+  X_{\m w} \f1M! \equiv \sum_{k\in\D_N\p{M}} w_k X_k,
   \label{eq:pert}
@@
   \D_N\p{M} \equiv
-  \set{ k \in \ZZ_N^M:~\t{all elements of}~k~\t{are distinct} },
+  \set{ k \in \ZZ_N^M: \t{all entries of}~k~\t{are distinct} },
   \label{eq:off_diags}
 \end{align}
-is the set of all ordered choices of $M$ spins out of $N$.  The
-normalization factor of $1/M!$ in \eqref{eq:pert} is included to
-account for the fact that there are $M!$ terms addressing any given
-choice of $M$ of spins.  Due to the permutational symmetry of the
-ground-state manifold $\M_0$ of the interaction Hamiltonian $H_0$, we
-will find it convenient to decompose
-$\D_N\p{M}=\pi\p{M}\times\C_N\p{M}$, where $\pi\p{M}$ is the
-permutation group of $\ZZ_M$, and $\C_N\p{M}$ is the set of all
-choices of $M$ spins out of $N$.  We can then expand
+is the strictly off-diagonal part of $\ZZ_N^M$.  The factor of $1/M!$
+is included in \eqref{eq:pert} by convention to average over the $M!$
+terms addressing any given set of spins.  Due to the permutational
+symmetry enforced by the interaction Hamiltonian $H_0$, we will find
+it convenient to decompose $\D_N\p{M}=\pi\p{M}\times\C_N\p{M}$, where
+$\pi\p{M}$ is the permutation group of $\ZZ_M$, and
+\begin{align}
+  \C_N\p{M} \equiv \set{ k \subset \ZZ_N: \abs{k} = M },
+  \label{eq:choices}
+\end{align}
+is the set of subsets (``choices'') of $M$ elements from $\ZZ_N$.  We
+can then expand
 \begin{align}
@@
 \end{align}
-For simplicity, we will assume throughout this work that $\O$ is
+For simplicity, we will assume throughout this work that $\hat\O$ is
 homogeneous in the sense that $M$ is the same for all
 $\p{X,\m w}\in\O$, with the understanding that a generalization to
-different $M$ for each $\p{X,\m w}\in\O$ is straightforward.
+inhomogeneous operators $\hat\O$ is straightforward.
 
@@
 \end{align}
-where $\EQPS$ denotes equality under a restriction to $\M_0$; $\bar w$
-is an average of $w_k$ over all $k\in\D_N\p{M}$; and $\col{X}$ is a
-permutation-symmetrized collective version of $X$, i.e.
-\begin{align}
-  \bar w \equiv \f1{\abs{\D_N\p{M}}} \sum_{k\in\D_N\p{M}} w_k,
+where $\EQPS$ denotes equality under a restriction to $\M_0$; the
+coefficient $\bar w$ is the average of all $w_k$; and the collective
+operator $\col{X}$ is a permutation-averaged sum of all $X_k$, i.e.
+\begin{align}
+  \bar w \equiv \f1{\abs{\D_N\p{M}}}
+  \sum_{k\in\D_N\p{M}} w_{\sigma\p{k}},
   &&
@@
 \end{align}
-where, to be explicit,
-\begin{align}
-  \abs{\D_N\p{M}} = N!_M,
-  &&
-  \pi\p{M} = M!,
-\end{align}
-with $N!_M\equiv\prod_{j=0}^{M-1}\p{N-j}=M!\times{N\choose M}$
-denoting a falling factorial.  An immediate consequence of
+with $\abs{\D_N\p{M}}=M!\times{N\choose M}=\prod_{j=0}^{M-1}\p{N-j}$
+and $\abs{\pi\p{M}}=M!$.  An immediate consequence of
 \eqref{eq:H_eff_1} is the fact that perturbing SU(2)-symmetric
@@
 \begin{align}
-  X_{\m w} \equiv \f1{M!} \sum_{k\in\D_N\p{M}} w_k X_k,
-  \label{eq:pert_copy}
+  X_{\m w}
+  \equiv \sum_{k\in\D_N\p{M}} w_k X_k
+  = \sum_{k\in\C_N\p{M}} \sum_{\sigma\in\pi\p{M}}
+  w_{\sigma\p{k}} X_{\sigma\p{k}},
 \end{align}
 where $X$ is an $M$-spin operator; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is
-a list of $M$ spins that $X_k$ acts on; $\D_N\p{M}$ is defined in
-\eqref{eq:off_diags}; and $\m w$ is a dimension-$M$ (i.e.~$M$-index)
-tensor with scalar entries $w_k$.
+a list of $M$ spins that $X_k$ acts on; $\m w$ is a dimension-$M$
+(i.e.~$M$-index) tensor with scalar entries $w_k$; $\D_N\p{M}$ is the
+strictly off-diagonal part of $\ZZ_N^M$, defined in
+\eqref{eq:off_diags}; $\C_N\p{M}$ is the set of all choices of $M$
+elements from $\ZZ_N$, defined in \eqref{eq:choices}; and $\pi\p{M}$
+is the permutation group of $\ZZ_M$.
 
@@
   \sum_{k_\alpha\in\D_N\p{M_\alpha}}
+  w_\alpha\p{k_\alpha} X_\alpha\p{k_\alpha}
+  = \f1{\m M!} \sum_{k\in\D_N\p{\m M}} \prod_{\alpha\in\ZZ_p}
   w_\alpha\p{k_\alpha} X_\alpha\p{k_\alpha},
@@
 $\m w\equiv\p{w_1,w_2,\cdots,w_p}$ is a vector of dimension-$M_\alpha$
-tensors $w_\alpha$;
-$k_\alpha\equiv\p{k_{\alpha,1},k_{\alpha,2},\cdots,k_{\alpha,M_\alpha}}$
-is a choice of $M_\alpha$ spins that $X_\alpha\p{k_\alpha}$ acts on;
-$\D_N\p{M_\alpha}$ is the set of all ordered choices of $M_\alpha$
-elements from a set of size $N$; and $w_\alpha\p{k_\alpha}$ is a
-scalar entry of $w_\alpha$.  For convenience, we define
-$w_\alpha\p{k_\alpha}$ and $X_\alpha\p{k_\alpha}$ for all
-$k_\alpha\in\ZZ_N^{M_\alpha}$ by enforcing that
-$w_\alpha\p{k_\alpha}=X_\alpha\p{k_\alpha}=0$ for all $k_\alpha$ with
-repeated values.  These definitions allow us to replace sums over
-$\D_N\p{M_\alpha}$ by sums over $\ZZ_N^{M_\alpha}$, thereby making the
-sums over the individual indices in $k_\alpha$ independent.
-
-To simplify the product in \eqref{eq:sym_prod_start}, we first replace
-all $\D_N\p{M_\alpha}$ by $\ZZ_N^{M_\alpha}$ and rewrite the product
-of sums as a sum of products:
-\begin{align}
-  \m X_{\m w}
-  &= \f1{M_1!} \sum_{k_1\in\ZZ_N^{M_1}}
-  \f1{M_2!} \sum_{k_2\in\ZZ_N^{M_1}} \cdots
-  \f1{M_p!} \sum_{k_p\in\ZZ_N^{M_p}}
-  \prod_{\alpha\in\ZZ_p} w_\alpha\p{k_\alpha} X_\alpha\p{k_\alpha} \\
-  &= \f1{\m M!} \sum_{k\in\ZZ_N^{\m M}} \prod_{\alpha\in\ZZ_p}
-  w_\alpha\p{k_\alpha} X_\alpha\p{k_\alpha},
-  \label{eq:sum_of_products}
-\end{align}
-where $\m M\equiv\p{M_1,M_2,\cdots,M_p}$ is a vector of the dimensions
-$M_\alpha$; $\m M!\equiv\prod_{M_\alpha\in\m M}M_\alpha!$ for
-shorthand; and
-$\ZZ_N^{\m M}\equiv\bigotimes_{\alpha\in\ZZ_p}\ZZ_N^{M_\alpha}$, such
-that each $k\in\ZZ_N^{\m M}$ has the decomposition
-$k=\p{k_1,k_2,\cdots,k_p}$ with $k_\alpha\in\ZZ_N^{M_\alpha}$.
-
-We now classify terms in \eqref{eq:sum_of_products} by the numbers
+tensors $w_\alpha$; $\m M\equiv\p{M_1,M_2,\cdots,M_p}$ is a vector of
+the dimensions $M_\alpha$;
+$k_\alpha=\p{k_{\alpha,1},k_{\alpha,2},\cdots,k_{\alpha,M_\alpha}}$ is
+a choice of $M_\alpha$ spins that $X_\alpha\p{k_\alpha}$ acts on;
+$w_\alpha\p{k_\alpha}$ is a scalar entry of $w_\alpha$;
+$\D_N\p{M_\alpha}$ is the strictly off-diagonal part of $\ZZ_N^M$,
+defined in \eqref{eq:off_diags};
+$\D_N\p{\m M}\equiv\bigotimes_{M_\alpha\in\m M}\D_N\p{M_\alpha}$, such
+that each $k\in\D_N\p{\m M}$ has the decomposition
+$k=\p{k_1,k_2,\cdots,k_p}$ with $k_\alpha\in\D_N\p{M_\alpha}$; and
+$\m M!\equiv\prod_{M_\alpha\in\m M}M_\alpha!$ for shorthand.  For
+later convenience, we define
+$w_\alpha\p{k_\alpha}=X_\alpha\p{k_\alpha}=0$ for any $k_\alpha$ with
+repeated entries.
+
+We now classify terms in \eqref{eq:sym_prod_start} by the numbers
 $g_S$ of indices shared by all tensors $w_\alpha$ with
@@
 \end{align}
-for all $\alpha\in\ZZ_p$.
-
-Schematically, a classification of the terms in
-\eqref{eq:sum_of_products} by diagrams $g\in\G\p{\m M}$ will allow us
-to expand
+for all $\alpha\in\ZZ_p$.  A classification of the terms in
+\eqref{eq:sym_prod_start} by diagrams $g\in\G\p{\m M}$ allows us to
+expand
 \begin{align}
@@
 \end{align}
-where $\EQPS$ denotes equality up to a restriction to the PS manifold
-$\M_0$; and $\m X_{\m w}\p{g}$ is the sum of all terms in
-\eqref{eq:sum_of_products} that are associated with the diagram $g$.
-To collect these terms, we first consider all distinct ways to assign
-$\abs{g}$ indices to specific tensor factors of $\m w$ and $\m X$ in a
-manner consistent with $g$.  For example, if $\m X=\p{X_1,X_2}$ with
-dimensions $\p{M_1,M_2}=\p{2,1}$ and $\abs{g}=2$ with
-$g_{\set{1,2}}=g_{\set{2}}=1$, then $X_1\p{a,b} X_2\p{b}$ and
-$X_1\p{b,a} X_2\p{b}$ are distinct assignments of $\abs{g}$ indices to
-$\m X$ in a manner consistent with $g$, provided that $a\ne b$.  After
-assigning $\abs{g}$ indices to $\m w$ and $\m X$ in a manner
-consistent with $g$, we need to sum over all values that these indices
-can take, which amounts to a sum over all
-$k=\p{k_1,k_2,\cdots,k_{\abs{g}}}\in\ZZ_N^{\abs{g}}$ with the
-restriction that $k_\mu\ne k_\nu$ for all $\mu\ne\nu$ to enforce that
-all indices are distinct, or equivalently a sum over all
-$k\in\D_N\p{\abs{g}}$.
+where $\EQPS$ denotes equality under a restriction to $\M_0$, and
+$\m X_{\m w}\p{g}$ is the sum of all terms in
+\eqref{eq:sym_prod_start} that are associated with the diagram $g$.
+
+To find an explicit form of $\m X_{\m w}\p{g}$, we first consider all
+distinct ways to assign $\abs{g}$ indices to specific tensor factors
+of $\m w$ and $\m X$ in a manner consistent with $g$.  For example, if
+$\m X=\p{X_1,X_2}$ with dimensions $\p{M_1,M_2}=\p{2,1}$ and
+$\abs{g}=2$ with $g_{\set{1,2}}=g_{\set{2}}=1$, then
+$X_1\p{a,b} X_2\p{b}$ and $X_1\p{b,a} X_2\p{b}$ are distinct
+assignments of $\abs{g}$ indices to $\m X$ in a manner consistent with
+$g$, provided that $a\ne b$.  After assigning $\abs{g}$ indices to
+$\m w$ and $\m X$ in a manner consistent with $g$, we need to sum over
+all values that these indices can take, which amounts to a sum over
+all $k\in\D_N\p{\abs{g}}$.
 
@@
 The assignment $X_1\p{a,b,c} X_2\p{a,b}$, for example, is the same as
-$X_1\p{b,a,c} X_2\p{b,a}$, because $a$ and $b$ are dummy indices.  In
-general, the treatment of the $g_S$ indistinguishable dots within
-region $S\in\PP^*\p{\ZZ_p}$ of $g$ as distinguishable results in
+$X_1\p{b,a,c} X_2\p{b,a}$, because $a$ and $b$ are dummy indices.
+Treating $g_S$ dummy indices as as distinguishable results in
 over-counting distinct index assignments by a factor of $g_S!$.
@@
 \begin{align}
-  \abs{\L\p{g}} = \f{\m M!}{\S\p{g}},
-  &&
-  \S\p{g} \equiv \prod_{S\in\PP^*\p{\ZZ_p}} g_S!,
-\end{align}
-where $\S\p{g}$ is a symmetry factor associated with the diagram $g$.
+  \abs{\L\p{g}} = \m M! \times \S\p{g},
+  &&
+  \S\p{g} \equiv \sp{\prod_{S\in\PP^*\p{\ZZ_p}} g_S!}^{-1}.
+  \label{eq:assignment_num}
+\end{align}
 Altogether, we have
 \begin{align}
-  \m X_{\m w}\p{g} \equiv \f1{\m M!} \sum_{\ell\in\L\p{g}}
-  \sum_{k\in\D_N\p{\abs{g}}} \m w\p{k,\ell} \m X\p{k,\ell},
-\end{align}
-where $\m w\p{k,\ell}$ and $\m X\p{k,\ell}$ denote the assignment of
-indices $k$ to all $w_\alpha$ and $X_\alpha$ according to $\ell$.  The
-specific choice of spins $k$ in $\m X\p{k,\ell}$ does not matter after
-projection onto the PS manifold $\M_0$, so the dependence of
-$\m X\p{k,\ell}$ on $k$ is superfluous, which is to say that we can
-write
-\begin{align}
-  \m X_{\m w}\p{g} \EQPS \f1{\m M!} \sum_{\ell\in\L\p{g}}
-  \sum_{k\in\D_N\p{\abs{g}}} \m w\p{k,\ell} \m X\p{\ell},
-\end{align}
-where $\EQPS$ denotes equality under a restriction to $\M_0$.  We now
-define, for any index assignment $\ell\in\L\p{g}$, the scalar
-\begin{align}
-  \m w\p{\ell}
-  \equiv \f1{\S\p{g}} \sum_{k\in\D_N\p{\abs{g}}} \m w\p{k,\ell},
-  \label{eq:assignment_scalar}
-\end{align}
-where the factor $1/\S\p{g}$ is included to average over the distinct
-ways to permute the shared indices in $\set{w_\alpha:\alpha\in S}$ for
-all $S\in\PP^*\p{\ZZ_p}$, which is essentially an average over the
-symmetries of $g$.  The motivation for this averaging comes from the
-consideration of the special case in which all $w_\alpha$ are
-invariant under arbitrary permutations of their tensor factors (or
-simply ``permutationally symmetric''),
-i.e.~$w_\alpha\p{k}=w_\alpha\p{\sigma\p{k}}$ for all permutations
-$\sigma\in\pi\p{M_\alpha}$.  In this case, we can simplify
-\begin{align}
-  \m w\p{\ell} = \sum_{k\in\C_N\p{g}} \m w\p{k,\ell},
-\end{align}
-where
-\begin{align}
-  \C_N\p{g} \equiv
-  \set{ k \in \bigotimes_{S\in\PP^*\p{\ZZ_p}} \C_N\p{g_S}
-    : \t{all elements of $k$ are distinct} }
-\end{align}
-is essentially the set of all index values $j,k\in\C_N\p{g}$ for which
-generally $\m w\p{j,\ell}\ne\m w\p{k,\ell}$ if $j\ne k$.  Note that if
-all $w_\alpha$ are permutationally symmetric, then $\m w\p{\ell}$
-depends on the diagram $g$, but not on the specific choice of index
-assignment $\ell\in\L\p{g}$.
-
-The definition of $\m w\p{\ell}$ in \eqref{eq:assignment_scalar}
-allows us to expand
-\begin{align}
-  \m X_{\m w}\p{g} \EQPS \f1{\abs{\L\p{g}}}
-  \sum_{\ell\in\L\p{g}} \m w\p{\ell} \m X\p{\ell},
+  \m X_{\m w}\p{g} \equiv \f{\S\p{g}}{\abs{\L\p{g}}}
+  \sum_{\ell\in\L\p{g}} \sum_{k\in\D_N\p{\abs{g}}}
+  \m w\p{k,\ell} \m X\p{k,\ell},
+\end{align}
+where $\m w\p{k,\ell}$ and $\m X\p{k,\ell}$ denote the scalar and
+operator acquired by assigning the indices
+$k=\p{k_1,k_2,\cdots,k_{\abs{g}}}$ to $\m w$ and $\m X$ according to
+$\ell$.  The specific choice of spins $k$ in $\m X\p{k,\ell}$ does not
+matter after projection onto the PS manifold $\M_0$, so the dependence
+of $\m X\p{k,\ell}$ on $k$ is superfluous, which allows us to write
+\begin{align}
+  \m X_{\m w}\p{g}
+  \EQPS \f{\S\p{g}}{\abs{\L\p{g}}} \sum_{\ell\in\L\p{g}}
+  \m w\p{\ell} \m X\p{\ell},
+  &&
+  \m w\p{\ell} \equiv \sum_{k\in\D_N\p{\abs{g}}} \m w\p{k,\ell}.
   \label{eq:diagram_terms}
 \end{align}
-where if additionally all tensors in $\m w$ {\it or} all operators in
-$\m X$ are permutationally symmetric, then
-\begin{align}
-  \m X_{\m w}\p{g} \EQPS \m w\p{g} \m X\p{g},
-  \label{eq:diagram_terms_perm}
-\end{align}
-with
+If, furthermore, all tensors in $\m w$ (or all operators in $\m X$)
+are {\it permutationally symmetric}, or invariant under arbitrary
+permutations of their tensor factors, then
+\begin{align}
+  \m X_{\m w}\p{g} \stackrel{\t{sym}}{=}_{\t{PS}}
+  \S\p{g} \m w\p{g} \m X\p{g},
+\end{align}
+where $\stackrel{\t{sym}}{=}_{\t{PS}}$ denotes equality up to
+\begin{enumerate*}
+\item permutational symmetry of all $\m w$ or $\m X$, and
+\item restriction to the PS manifold $\M_0$; and
+\end{enumerate*}
 \begin{align}
@@
   \sum_{\ell\in\L\p{g}} \m X\p{\ell}.
-  \label{eq:assignment_averages}
-\end{align}
-The simplification in \eqref{eq:diagram_terms_perm} also holds if all
-$X_\alpha$ are permutationally symmetric, in which case $\m w\p{g}$ is
-similarly defined as an average of $\m w\p{\ell}$ over all
-$\ell\in\L\p{g}$.
-
+  \label{eq:diagram_factors}
+\end{align}
 In summary, the product of operators $\m X_{\m w}$ in
@@
 \eqref{eq:sym_prod_diagrams}, as a sum of $\m X_{\m w}\p{g}$ over all
-diagrams $g\in\G\p{\m M}$.  Each term $\m X_{\m w}\p{g}$ is
-essentially an average of $\m w\p{\ell} \m X\p{\ell}$ over all
-distinct index assignments $\ell\in\L\p{g}$, as in
+diagrams $g\in\G\p{\m M}$.  Each term $\m X_{\m w}\p{g}$ is, up to a
+symmetry factor $\S\p{g}$, an average of $\m w\p{\ell} \m X\p{\ell}$
+over all distinct index assignments $\ell\in\L\p{g}$, as in
 \eqref{eq:diagram_terms}.  The scalar $\m w\p{\ell}$ is obtained by
@@
 $\m w$ as prescribed by the $\ell$, and summing over all
-$k\in\D_N\p{\abs{g}}$, as in \eqref{eq:assignment_scalar}.  The
-$\abs{g}$-spin operator $\m X\p{\ell}$ is obtained by contracting the
-$M_\alpha$-spin operators $X_\alpha$ in $\m X$ in accordance with the
-index assignment $\ell$.  If the tensors $\m w$ or operators $\m X$
-are permutationally symmetric, then $\m X_{\m w}\p{g}$ factorizes into
-$\m w\p{g} \m X\p{g}$, where $\m w\p{g}$ and $\m X\p{g}$ are
-respectively defined as the average of $\m w\p{\ell}$ and
-$\m X\p{\ell}$ over all $\ell\in\L\p{g}$.
+$k\in\D_N\p{\abs{g}}$, while the $\abs{g}$-spin operator
+$\m X\p{\ell}$ is obtained by contracting the operators in $\m X$ in
+accordance with the index assignment $\ell$.  If the tensors $\m w$ or
+operators $\m X$ are all permutationally symmetric, then
+$\m X_{\m w}\p{g}$ factorizes into $\S\p{g} \m w\p{g} \m X\p{g}$,
+where $\m w\p{g}$ and $\m X\p{g}$ are respectively averages of
+$\m w\p{\ell}$ and $\m X\p{\ell}$ over all index assignments
+$\ell\in\L\p{g}$.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Formalizing diagrams}
+\subsection{Formalizing diagrams and minimizing computational costs}
 \label{sec:diagrams}
@@
 simplifying products of multi-body operators, through the expansion in
-\eqref{eq:sym_prod_diagrams}.  Here, we define these diagrams as
-formal objects that enable us to systematically perform otherwise
-intractable calculations.  We then derive rules to ``reduce'' these
-diagrams in such a way as to minimize the computational cost of their
-numerical evaluation.
-
-All diagrams consist of $p$ circles, and label each circle
-$\alpha\in\ZZ_p$ by some tensor $w_\alpha$, as in
-\eqref{eq:venn_diagram} (or below).  The simplest diagrams have $g_S$
-filled dots (``$\bullet$'') at the intersection of the circles
-$\alpha\in S\subset\ZZ_p$, and are simply equal to the scalar
-$\m w\p{g}$ defined in \eqref{eq:assignment_averages}.  If all tensors
-$w_\alpha$ are permutationally symmetric, for example, then
+\eqref{eq:sym_prod_diagrams}.  The purpose of diagrams is to
+succinctly represent the scalar content of each term in this
+expansion.  Here, we define these diagrams as formal objects that
+enable us to systematically perform otherwise intractable
+calculations.  We then derive rules to ``reduce'' these diagrams in
+such a way as to minimize the computational cost of their numerical
+evaluation.
+
+All diagrams consist of $p$ circles, with each circle $\alpha\in\ZZ_p$
+labeled by some tensor $w_\alpha$, as in \eqref{eq:venn_diagram} (or
+below).  The simplest diagrams have $g_S$ filled dots (``$\bullet$'')
+at the intersection of the circles $\alpha\in S\subset\ZZ_p$, and are
+simply equal to the scalar $\S\p{g} \m w\p{g}$ defined by
+\eqref{eq:assignment_num}, \eqref{eq:diagram_terms}, and
+\eqref{eq:diagram_factors}.  If all tensors $w_\alpha$ are
+permutationally symmetric, for example, then
 \begin{align}
   \diagram{example_123}
-  = \sum_{\substack{a,b,c,\in\ZZ_N\\\p{d,e}\in\C_N\p{2}\\
-      \abs{\set{a,b,c,d,e}}=5}}
-  w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e},
+  = \f1{2!} \sum_{\p{a,b,c,d,e}\in\D_N\p{5}}
+  w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}.
   \label{eq:example_123}
 \end{align}
-where the restriction $\abs{\set{a,b,c,d,e}}=5$ denotes that all of
-$a,b,c,d,e$ are distinct.  When the tensors $w_\alpha$ are not all
-permutationally symmetric, one should generally equip each diagram
-with an index assignment $\ell\in\L\p{g}$, in which case the value of
-this diagram is equal to the scalar $\m w\p{\ell}$ defined in
-\eqref{eq:assignment_scalar}; if no such assignment is provided, then
-the diagram is equal to the average over all index assignments
-consistent with $g$.  To keep our discussions simple, we will continue
-to assume that all tensors $w_\alpha$ are permutationally symmetric,
-but note that all discussions generally apply to the asymmetric case
-as well, as long as all sums over $\C_N\p{M}$ are replaced by a sums
-over $\D_N\p{M}$ with a symmetry prefactor of $1/M!$,
-i.e.~$\sum_{k\in\C_N\p{M}}f\p{k} \to
-\p{1/M!}\sum_{k\in\D_N\p{M}}f\p{k}$.  The only exception, in which
-diagrams equipped with index assignments $\ell\in\L\p{g}$ are treated
-differently than diagrams without such assignments, concerns one step
-in the process of ``reducing'' diagrams (specifically, the step of
-``eliminating crosses'').  We will comment on this exception when it
-appears.
+When the tensors $\m w$ are not all permutationally symmetric, one can
+additionally equip a diagram with a specific index assignment
+$\ell\in\L\p{g}$, in which case the diagram is equal to
+$\S\p{g} \m w\p{\ell}$.  Diagrams that are not labeled with a specific
+index assignment $\ell$ are then an average over all $\ell\in\L\p{g}$.
+To keep mathematical expressions simple, we will use symmetric tensors
+in all examples in this section, but note that our discussions apply
+just as well to the asymmetric case.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Reducing computational costs}
+\subsubsection{Reducing diagrams}
 
@@
 
-If a region of a diagram contains $F$ filled dots, those dots
-correspond to indices that sum over $\C_N\p{F}$ with the constraint
-that they are not equal to any other filled dots in a diagram.  $E$
-empty dots in a region, meanwhile, correspond to indices that are
-summed over $\C_N\p{E}$ without any additional constraints, e.g.
+Filled dots in a diagram correspond to indices that are constrained to
+have different values.  The presence of $F$ filled dots thus
+corresponds to a sum over $\D_N\p{F}$.  Every empty dot, meanwhile,
+corresponds to an index that is summed over $\ZZ_N$ without additional
+constraints, e.g.
 \begin{align}
   \diagram{example_o}
-  = \sum_{\substack{\p{a,b}\in\C_N\p{2}\\e\in\ZZ_N\\
-      \abs{\set{a,b,e}}=3}}
-  \sum_{\substack{\p{c,d}\in\C_N\p{2}\\f\in\ZZ_N}}
-  v\p{a,b,c,d,e} w\p{e,f}.
-\end{align}
-Empty dots are therefore entirely independent of filled dots.
-Crosses, meanwhile, are {\it entirely dependent} on filled dots:
-``cross indices'' are summed {\it only} over the values of ``filled
-dot indices'', e.g.
+  \propto \sum_{\p{a,b,e}\in\D_N\p{3}} \sum_{c,d,e\in\ZZ_N}
+  v\p{a,b,c,d,e} w\p{e,f},
+\end{align}
+where the proportionality is up to an appropriate symmetry factor (in
+this case, $1/2!\times1/2!$; clarified in Section
+\ref{sec:symmetry_factors}).  Crosses, meanwhile, correspond to
+indices that are {\it only} summed over the values of ``filled dot
+indices'', e.g.
 \begin{align}
   \diagram{example_x}
-  = \sum_{\substack{a,e\in\ZZ_N\\\abs{\set{a,e}}=2}}
-  \sum_{b,c\in\set{a,e}} \sum_{d\in\ZZ_N}
-  u\p{a,b,c} v\p{b,d,e} w\p{b,c,e}.
+  = \sum_{\p{a,e}\in\D_N\p{2}} \sum_{b,c\in\set{a,e}}
+  \sum_{d\in\ZZ_N} u\p{a,b,c} v\p{b,d,e} w\p{b,c,e}.
 \end{align}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Decomposing diagrams}
-
 Given a diagram containing only dots (filled or empty), we can
@@
   \diagram{example_elim}
-  \equiv \sum_{\substack{a,b,c\in\ZZ_N\\\abs{\set{a,b,c}}=3}}
-  v\p{a,b} w\p{b,c},
+  = \sum_{\p{a,b,c}\in\D_N\p{3}} v\p{a,b} w\p{b,c},
 \end{align}
@@
   \diagram{example_elim_o}
-  \equiv \sum_{\substack{b,c\in\ZZ_N\\\abs{\set{b,c}}=2}}
-  v\p{\circ,b} w\p{b,c},
+  = \sum_{\p{b,c}\in\D_N\p{2}} v\p{\circ,b} w\p{b,c},
   &&
@@
   \diagram{example_elim_x}
-  \equiv \sum_{\substack{b,c\in\ZZ_N\\\abs{\set{b,c}}=2}}
-  \sum_{a\in\set{b,c}} v\p{a,b} w\p{b,c}.
+  = \sum_{\p{b,c}\in\D_N\p{2}} \sum_{a\in\set{b,c}}
+  v\p{a,b} w\p{b,c}.
+  \label{eq:example_elim_x}
 \end{align}
@@
 a diagram, we first note that the corresponding index can ignore other
-indices on tensors that contain the index; that is (pay attention to
-the sum over $a$),
+indices on tensors that contain the index, which is to say that we can
+replace the sum over $a\in\set{b,c}$ in \eqref{eq:example_elim_x} by a
+sum over $a\in\set{c}$.  For each remaining index addressed by a
+cross, meanwhile, the cross simply forces the corresponding indices to
+be equal, which is equivalent to moving a filled dot into a different
+region:
 \begin{align}
   \diagram{example_elim_x}
-  \equiv \sum_{\substack{b,c\in\ZZ_N\\\abs{\set{b,c}}=2}}
-  \sum_{a\in\set{b,c}} v\p{a,b} w\p{b,c}
-  = \sum_{\substack{b,c\in\ZZ_N\\\abs{\set{b,c}}=2}}
-  \sum_{a\in\set{c}} v\p{a,b} w\p{b,c}.
-\end{align}
-For each remaining index addressed by a cross, meanwhile, the cross
-simply forces the corresponding indices to be equal, which is
-equivalent to moving a filled dot into a different region:
-\begin{align}
-  \diagram{example_elim_x}
-  = 2 \diagram{example_elim_x_full}
-  = 2 \sum_{\p{b,c}\in\C_N\p{2}} v\p{c,b} w\p{b,c}.
+  = \sum_{\p{b,c}\in\D_N\p{2}} \sum_{a\in\set{c}} v\p{a,b} w\p{b,c}
+  = \sum_{\p{b,c}\in\D_N\p{2}} v\p{c,b} w\p{b,c}
+  = 2 \diagram{example_elim_x_full}.
   \label{eq:example_elim_final}
 \end{align}
-The factor of 2 accounts for the fact that the left diagram sums over
-$b,c\in\ZZ_N$, while the right diagram sums over
-$\p{b,c}\in\C_N\p{2}$.  We can generally repeat the above process of
-eliminating filled dots and crosses until only empty dots remain in a
-diagram.  The only complication in carrying out this process is that
-we have to manually keep track of symmetry factors at each step, such
-as the factor of 2 in \eqref{eq:example_elim_final}.
+The added factor of 2 accounts for the fact that the two diagrams in
+\eqref{eq:example_elim_final} have different symmetry factors, whereas
+re-writing index constraints (i.e.~by eliminating crosses) does not
+affect any numerical prefactors on the tensor contraction represented
+by a diagram.  We can generally repeat the above process of
+sequentially eliminating filled dots and crosses until only empty dots
+remain.  The only complication in carrying out this process is that,
+in the case of permutationally symmetric tensors, we have to manually
+keep track of symmetry factors such as the factor of 2 above.
 
@@
 \subsubsection{Symmetry factors}
-\label{sec:symmertry_factors}
+\label{sec:symmetry_factors}
 
 As seen in \eqref{eq:example_elim_final}, the process of simplifying
-diagrams generally involves symmetry factors that need to be kept
-track of by hand.  Here we discuss the rules governing symmetry
-factors that appear when decomposing diagrams.  In general,
-eliminating a filled dot from a region with $F$ filled dots makes a
-diagram pick up a factor of $1/F$, e.g.
+diagrams with permutationally symmetric tensors generally involves
+symmetry factors that need to be kept track of by hand.  Here we
+discuss the rules governing symmetry factors that appear when
+decomposing such diagrams.  In general, eliminating a filled dot from
+a region with $F$ filled dots makes a diagram pick up a factor of
+$1/F$, e.g.
 \begin{align}
@@
 \end{align}
-This factor simply accounts for the fact that the diagram on the left
-has an overall symmetry factor of $1/4!\times1/4!$, while the diagrams
-on the right only have an overall symmetry factor of $1/3!\times1/4!$.
 Similarly, adding an empty dot to a region with $E$ empty dots makes
@@
 \begin{align}
-  \diagram{example_sym_x}
-  = 2 \diagram{example_sym_x_elim},
+  \diagram{example_sym_x} = 2 \diagram{example_sym_x_elim},
 \end{align}
@@
 $F_{\t{old}}$ choices of which dot to move.  More precisely,
-eliminating a cross and moving a dot from the old region to the new
-one corresponds to enforcing that the index associated with the cross
-take on the same value as an index associated with a filled dot in the
-old region, and there are $F_{\t{old}}$ such indices to choose from.
-
-If the tensors $\set{w_\alpha:\alpha\in\ZZ_p}$ associated with a
-diagram are not symmetric, then all dots in a given region
-$S\subset\ZZ_p$ of a diagram are distinguishable by their association
-with a specific tensor factor of each $w_\alpha$ with $\alpha\in S$.
-Rather than picking up a factor of $F_{\t{old}}$ to account for the
-number of ways to move a dot from the old region to the new one, one
-should then sum over each choice of which dot to move, e.g.
-\begin{align}
-  \diagram{example_sym_x_color}
-  = \f13 \sp{\diagram{example_sym_x_color_r}
-    + \diagram{example_sym_x_color_m}
-    + \diagram{example_sym_x_color_b}}.
-\end{align}
+eliminating a cross and moving a dot corresponds to forcing the index
+associated with the cross to be equal to the index associated with the
+filled dot (in the ``old'' region), and there are $F_{\t{old}}$ such
+indices to choose from.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Matrix elements of individual multi-spin operators}
-
-The process of constructing and simplifying diagrams to compute the
-coefficients $\m w\p{g}$ of the diagrammatic operator product
-expansion in \eqref{eq:sym_prod_diagrams} is systematic enough to be
-converted into an algorithm for execution on a computer.  The last
-step necessary to automate the evaluation of
-\eqref{eq:sym_prod_diagrams} is expansion of the multi-spin operators
-$\m X\p{g}$ in a basis for the PS manifold $\M_0$.  Without loss of
-generality, we consider a single permutationally symmetric $M$-spin
-operator $X$; our task is essentially to find coefficients of the
-expansion
+\subsection{Matrix elements of multi-spin operators}
+
+The process of constructing and simplifying diagrams to compute
+coefficients of the diagrammatic operator product expansion in
+\eqref{eq:sym_prod_diagrams} is systematic enough to be converted into
+an algorithm for execution on a computer.  The last step necessary to
+automate the evaluation of \eqref{eq:sym_prod_diagrams} is the
+expansion of the multi-spin operators $\m X\p{\ell}$ in a basis for
+the PS manifold $\M_0$.  Without loss of generality, we consider a
+single $M$-spin operator $X$; our task is essentially to find
+coefficients of the expansion
 \begin{align}
@@
 $N$ spins with $n$ states each, a suitable basis for $\M_0$ can be
-labeled by the occupation number of each single-spin state, namely
-$a=\p{a_1,a_2,\cdots,a_n}$ with each $a_\mu\ge 0$ and
-$\sum_\mu a_\mu=N$.  Written out explicitly,
+labeled by the occupation number $a_\mu$ of each single-spin state,
+namely $a=\p{a_1,a_2,\cdots,a_n}$ with $\sum_\mu a_\mu=N$.  Written
+out explicitly,
 \begin{align}
@@
 
-\red{[TODO: fix section to consider the case without permutational
-  symmetry]}
-
 Here we simplify the two-body product
@@
   X_{\m v} Y_{\m w}
-  = \sp{\sum_{\p{k,\ell},\p{p,q}\in\C_N\p{2}}
+  = \sp{\sum_{\p{k,\ell},\p{p,q}\in\D_N\p{2}}
     v_{k\ell} w_{pq} X_{k\ell} Y_{pq}} \\
-  \EQPS \diagram{two_body_0} X_{1,2} Y_{3,4}
-  + \diagram{two_body_1} \sp{X,Y}_{1,2,3}
-  + \diagram{two_body_2} \sp{X,Y}_{1,2},
+  \EQPS \diagram{two_body_0} X_{1,2} Y_{3,4} + \diagram{two_body_1}
+  X_{1,2} Y_{2,3} + \diagram{two_body_2} X_{1,2} Y_{1,2},
 \end{multline}
 where $\P_0$ is a projector onto the PS manifold $\M_0$; $k,\ell,p,q$
-index individual spins; $v_{k\ell},w_{pq}$ are scalar coefficients
-with $m_{pq}=m_{qp}$ and $m_{pp}=0$ for both $\m m\in\set{\m v,\m w}$;
-and $X,Y$ are permutationally symmetric two-spin operators; and
-\begin{align}
-  \sp{X,Y}_{1,2,3}
-  \equiv \f14 X_{1,2}\p{Y_{1,3} + Y_{2,3} + Y_{3,1} + Y_{3,2}},
-  &&
-  \sp{X,Y}_{1,2} \equiv \f12 X_{1,2} \p{Y_{1,2} + Y_{2,1}}.
-\end{align}
+index individual spins that the two-body operators $X_{k\ell},Y_{pq}$
+act on; and $v_{k\ell},w_{rs}$ are scalar coefficients with
+$m_{pq}=m_{qp}$ and $m_{pp}=0$ for each of $\m m\in\set{\m v,\m w}$.
 This product appears, for example, in the calculation of the
-second-order effective Hamiltonian $H_2^{(2)}$ induced on the
-permutationally symmetric manifold $\M_0$ by a two-body perturbation.
+second-order effective Hamiltonian $H_2^{(2)}$ induced on the PS
+manifold $\M_0$ by a permutationally symmetric two-body perturbation.
 Defining
 \begin{align}
-  \v{\m m} \equiv \sum_{\p{p,q}\in\C_N\p{2}} m_{pq} \ket{\set{p,q}},
+  \v{\m m} \equiv \sum_{\p{p,q}\in\C_N\p{2}} m_{pq} \ket{p,q},
   &&
@@
 \end{align}
-and simplify
+and
 \begin{align}
@@
   + \diagram{two_body_2_oo}
-  = \col{v}\,\col{w} -\v v\c\v w + \v{\m v}\c\v{\m w}.
+  = \col{v}\,\col{w} - \v v\c\v w + \v{\m v}\c\v{\m w}.
 \end{align}
@@
   X_{\m v} Y_{\m w}
-  \EQPS \p{\col{v}\,\col{w} -\v v\c\v w + \v{\m v}\c\v{\m w}}
+  \EQPS \p{\col{v}\,\col{w} - \v v\c\v w + \v{\m v}\c\v{\m w}}
   X_{1,2} Y_{3,4}
-  + \p{\v v\c\v w - 2 \v{\m v} \c \v{\m w}} \sp{X,Y}_{1,2,3}
-  + \v{\m v} \c \v{\m w}\, \sp{X,Y}_{1,2}.
+  + \p{\v v\c\v w - 2 \v{\m v} \c \v{\m w}} X_{1,2} Y_{2,3}
+  + \v{\m v} \c \v{\m w}\, X_{1,2} Y_{1,2}.
 \end{align}
@@
 \begin{align}
-  \sum_{\p{k,\ell},\p{p,q}\in\C_N\p{2}} v_{k\ell} w_{pq}
+  \sum_{\p{k,\ell},\p{p,q}\in\D_N\p{2}} v_{k\ell} w_{pq}
   Z_k Z_\ell Z_p Z_q
-  \EQPS \p{\col{v}\,\col{w} -\v v\c\v w + \v{\m v}\c\v{\m w}}
+  \EQPS \p{\col{v}\,\col{w} - \v v\c\v w + \v{\m v}\c\v{\m w}}
   Z^{\otimes 4}
-  + \p{\v v\c\v w - 2 \v{\m v} \c \v{\m w}} Z^{\otimes 2}
-  + \v{\m v}\c\v{\m w}.
-\end{align}
-In terms of collective operators
-$\col{Z} \equiv \sum_{p\in\ZZ_N} Z_p$, we then have
+  + \p{\v v\c\v w - \v{\m v} \c \v{\m w}} Z^{\otimes 2}
+  + \v{\m v}\c\v{\m w},
+\end{align}
+which we can write in terms of collective operators
+$\col{Z}\equiv\sum_{p\in\ZZ_N}Z_p$ as
 \begin{multline}
@@
   Z_k Z_\ell Z_p Z_q \\
-  \EQPS \f{\col{v}\,\col{w} -\v v\c\v w + \v{\m v}\c\v{\m w}}{N!_4}
+  \EQPS \f{\col{v}\,\col{w} - \v v\c\v w + \v{\m v}\c\v{\m w}}{N!_4}
   \sp{\col{Z}^4 - 2\p{3N-4} \col{Z}^2}
-  + \f{\p{\v v\c\v w - 2 \v{\m v} \c \v{\m w}}}{N!_2}\, \col{Z}^2 \\
+  + \f{\v v\c\v w - 2 \v{\m v} \c \v{\m w}}{N!_2}\, \col{Z}^2 \\
   + \f{3\,\col{v}\,\col{w} - N \v v\c\v w + N\p{N-2}\v{\m v}\c\v{\m
@@
 \subsection{Three two-body Ising-like operators}
-
-\red{[TODO: fix section to consider the case without permutational
-  symmetry]}
 

### 2020-03-25 collective_sun/scripts/make_diagrams.py

@@ -238,7 +238,7 @@ plt.savefig(fig_dir + "example_o.pdf")
 
 dots = { "a" : 1, "abc" : 1, "ac" : 1, "b" : 1, "bc" : 1 }
 markers = { "abc" : "x", "ac" : "x", "b" : "o" }
-make_triple_diagram(dots, markers, figsize = (1,1))
+make_triple_diagram(dots, markers, figsize = (0.8,0.8))
 add_triple_labels()
 plt.tight_layout(pad = 0)
 plt.savefig(fig_dir + "example_x.pdf")
@@ -257,11 +257,6 @@ make_double_diagram(dots, { "a" : "o" })
 add_double_labels()
 plt.tight_layout(pad = 0.15)
 plt.savefig(fig_dir + "example_elim_o.pdf")
-
-make_double_diagram(dots, { "a" : "x" })
-add_double_labels()
-plt.tight_layout(pad = 0.15)
-plt.savefig(fig_dir + "example_elim_x.pdf")
 
 make_double_diagram(dots, { "a" : "x" })
 add_double_labels()
@@ -309,36 +304,6 @@ make_double_diagram(dots)
 add_double_labels()
 plt.tight_layout(pad = 0.15)
 plt.savefig(fig_dir + "example_sym_x_elim.pdf")
-
-dots = { "a" : 2, "b" : 3 }
-make_double_diagram(dots, { "a" : "x" })
-add_double_labels()
-plt.tight_layout(pad = 0.15)
-plt.savefig(fig_dir + "example_sym_x_color.pdf")
-
-dots = { "a" : 2, "b" : ( 3, [ "r", "m", "b" ]) }
-make_double_diagram(dots, { "a" : "x" })
-add_double_labels()
-plt.tight_layout(pad = 0.15)
-plt.savefig(fig_dir + "example_sym_x_color.pdf")
-
-dots = { "a" : 1, "b" : ( 2, [ "m", "b" ]), "ab" : ( 1, ["r"] ) }
-make_double_diagram(dots)
-add_double_labels()
-plt.tight_layout(pad = 0.15)
-plt.savefig(fig_dir + "example_sym_x_color_r.pdf")
-
-dots = { "a" : 1, "b" : ( 2, [ "r", "b" ]), "ab" : ( 1, ["m"] ) }
-make_double_diagram(dots)
-add_double_labels()
-plt.tight_layout(pad = 0.15)
[…]

### 2020-03-25 collective_sun/sun_perturbation.tex
fix typo in notes

@@
   &&
-  X_{\m w} \f1M! \equiv \sum_{k\in\D_N\p{M}} w_k X_k,
+  X_{\m w} \equiv \f1{M!} \sum_{k\in\D_N\p{M}} w_k X_k,
   \label{eq:pert}
@@
   X_{\m w}
-  \equiv \sum_{k\in\D_N\p{M}} w_k X_k
-  = \sum_{k\in\C_N\p{M}} \sum_{\sigma\in\pi\p{M}}
+  \equiv \f1{M!} \sum_{k\in\D_N\p{M}} w_k X_k
+  = \f1{M!} \sum_{\sigma\in\pi\p{M}} \sum_{k\in\C_N\p{M}}
   w_{\sigma\p{k}} X_{\sigma\p{k}},
@@
 strictly off-diagonal part of $\ZZ_N^M$, defined in
-\eqref{eq:off_diags}; $\C_N\p{M}$ is the set of all choices of $M$
-elements from $\ZZ_N$, defined in \eqref{eq:choices}; and $\pi\p{M}$
-is the permutation group of $\ZZ_M$.
+\eqref{eq:off_diags}; $\pi\p{M}$ is the permutation group of $\ZZ_M$;
+and $\C_N\p{M}$ is the set of subsets (``choices'') of $M$ elements
+from $\ZZ_N$, defined in \eqref{eq:choices}.
 

### 2020-03-25 collective_sun/sun_perturbation.tex
fix section on the multi-body eigenvalue problem to account for multi-body operators without permutational symmetry some additional minor fixes elsewhere

@@
 
+\newcommand{\lra}{\leftrightarrow}
 \newcommand{\floor}[1]{\lfloor{#1}\rfloor}
-\newcommand{\ceil}[1]{\lceil{#1}\rceil}
 
@@
 \def\oop#1#2{\oket{#1}\!\obra{#2}}
-
-\usepackage{mathtools} % for coloneqq
 
@@
 
-\red{[TODO: fix section to consider the case without permutational
-  symmetry of $X_{\m w}$]}
-
 In this section, we simplify the products of the form
@@
 and $\C_N\p{M}$ is the set of subsets (``choices'') of $M$ elements
-from $\ZZ_N$, defined in \eqref{eq:choices}.
+from $\ZZ_N$, defined in \eqref{eq:choices}.  For shorthand, we will
+define $\tilde X_{\m w}\equiv M! X_{\m w}$.
 
@@
 \begin{align}
-  H_0 X_{\m w} \ket\psi
-  = \sum_{\substack{\p{p,q}\in\C_N\p{2}\\k\in\ZZ_N}}
+  H_0 \tilde X_{\m w} \ket\psi
+  = \sum_{\substack{k\in\ZZ_N\\\set{p,q}\in\C_N\p{2}}}
   h_{pq} w_k \Pi_{pq} X_k \ket\psi.
@@
 The sum in \eqref{eq:single_body_diagnosis_start} has terms with
-$k\in\set{p,q}$, and terms with $k\notin\set{p,q}$.  In the case of
+$k\notin\set{p,q}$, and terms with $k\in\set{p,q}$.  In the case of
 $k\notin\set{p,q}$, the permutation operator $\Pi_{pq}$ commutes with
-$X_k$ and annihilates on $\ket\psi$, leaving terms at each fixed $k$
-of the form
-\begin{align}
-  \sum_{\substack{\p{p,q}\in\C_N\p{2}\\k\notin\set{p,q}}} h_{pq}
-  = \sum_{\p{p,q}\in\C_N\p{2}} h_{pq}
-  - \sum_{\substack{\p{p,q}\in\C_N\p{2}\\k\in\set{p,q}}} h_{pq}
-  = E_0 - \f12 \sum_{\substack{p,q\in\ZZ_N\\k\in\set{p,q}}} h_{pq}
-  = E_0 - h_k,
-\end{align}
-where we define $h_{pq}$ for arbitrary $p,q$ by enforcing
-$h_{pq}=h_{qp}$ and $h_{pp}=0$; and $h_k \equiv \sum_\ell h_{k\ell}$.
+$X_k$ and annihilates on $\ket\psi$, so
+\begin{align}
+  \sum_{k\in\ZZ_N} \sum_{\substack{\set{p,q}\in\C_N\p{2}\\k\notin\set{p,q}}}
+  h_{pq} w_k X_k \ket\psi
+  = E_0 \tilde X_{\m w} \ket\psi
+  - \sum_{k\in\ZZ_N} \sum_{\substack{\set{p,q}\in\C_N\p{2}\\k\in\set{p,q}}}
+  h_{pq} w_k X_k \ket\psi,
+\end{align}
+where $E_0 = \sum_{\set{p,q}\in\C_N\p{2}} h_{pq}$ is the interaction
+energy of PS states $\ket\psi\in\M_0$.  Simplifying
+\begin{align}
+  \sum_{\substack{\set{p,q}\in\C_N\p{2}\\k\in\set{p,q}}} h_{pq}
+  = \f12 \sum_{\substack{p,q\in\ZZ_N\\k\in\set{p,q}}} h_{pq}
+  = \f12 \sum_{p\in\ZZ_N} h_{pk} + \f12 \sum_{q\in\ZZ_N} h_{kq}
+  = h_k,
+  &&
+  h_k \equiv \sum_{p\in\ZZ_N} h_{pk},
+\end{align}
+we can write
+\begin{align}
+  \sum_{k\in\ZZ_N} \sum_{\substack{\set{p,q}\in\C_N\p{2}\\k\notin\set{p,q}}}
+  h_{pq} w_k X_k \ket\psi
+  = E_0 \tilde X_{\m w} \ket\psi
+  - \sum_{k\in\ZZ_N} h_k w_k X_k \ket\psi.
+\end{align}
 The terms in \eqref{eq:single_body_diagnosis_start} with
-$k\notin\set{p,q}$ are then
-\begin{align}
-  \sum_{\substack{\p{p,q}\in\C_N\p{2}\\k\notin\set{p,q}}}
+$k\in\set{p,q}$, meanwhile, are
+\begin{align}
+  \sum_{\substack{\set{p,q}\in\C_N\p{2}\\k\in\set{p,q}}}
   h_{pq} w_k \Pi_{pq} X_k \ket\psi
-  = \sum_{k\in\ZZ_N} \p{E_0 - h_k} w_k X_k \ket\psi
-  = E_0 X_{\m w} \ket\psi - \sum_{k\in\ZZ_N} h_k w_k X_k \ket\psi
-\end{align}
-while the terms in \eqref{eq:single_body_diagnosis_start} with
-$k\in\set{p,q}$ take the form
-\begin{align}
-  \sum_{\substack{\p{p,q}\in\C_N\p{2}\\k\in\set{p,q}}}
-  h_{pq} v_k \Pi_{pq} X_k \ket\psi
-  = \sum_{\p{k,q}\in\C_N\p{2}} h_{kq} w_k X_q \ket\psi
-  + \sum_{\p{p,k}\in\C_N\p{2}} h_{pk} w_k X_p \ket\psi
-  = \sum_{k,q\in\ZZ_N} h_{kq} w_k X_q \ket\psi
-\end{align}
-which implies
+  = \sum_{k,q\in\ZZ_N} h_{kq} w_k X_q \ket\psi,
+\end{align}
+so in total
 \begin{align}
@@
 all couplings $h_{pq}$; the vector
-$\v h\equiv\sum_{p,q\in\ZZ_N} h_{pq} \ket{p}$ is the sum of all
-columns of $\m h$; and $\diag\v h$ is a matrix with $\v h$ on the
-diagonal and zeroes everywhere else.
+$\v h \equiv \sum_{p\in\ZZ_N} h_p \ket{p} = \sum_{p,q\in\ZZ_N} h_{pq}
+\ket{p}$ is the sum of all columns of $\m h$; and $\diag\v h$ is a
+matrix with $\v h$ on the diagonal and zeroes everywhere else.
 
@@
 solvable analytically.  In this case, the couplings $h_{pq}$ depend
-only on the separation $\abs{p-q}$, so the eigenvectors of $\m h$ are
-plane waves of the form
+only on the separation $\abs{p-q}$, so eigenvectors of $\m h$ are
+therefore plane waves of the form
 \begin{align}
@@
   = \sum_{p,d\in\ZZ_L^D} h_{p,p+d} e^{i\p{p+d}\c k} \ket{p}
-  = \sum_{d\in\ZZ_L^D} h_{0,d} e^{id\c k} \v{\m w}_k
   = \sum_{d\in\ZZ_L^D} h_{0,d} \cos\p{d\c k} \v{\m w}_k,
@@
 \begin{align}
-  \p{\m h - \diag\v h}\c\v v_k = \Delta_k \v{\m w}_k,
+  \p{\m h - \diag\v h}\c\v{\m w}_k = \Delta_k \v{\m w}_k,
   &&
@@
 
-\red{[TODO: fix section to consider the case without permutational
-  symmetry of $X_{\m w}$]}
-
 We now consider the full $M$-body case
 \begin{align}
-  H_0 X_{\m w} \ket\psi
-  = \sum_{\substack{\p{p,q}\in\C_N\p{2}\\k\in\C_N\p{M}}}
+  H_0 \tilde X_{\m w} \ket\psi
+  = \sum_{\substack{k\in\D_N\p{M}\\\set{p,q}\in\C_N\p{2}}}
   h_{pq} w_k \Pi_{pq} X_k \ket\psi.
@@
 \end{align}
-The sum in \eqref{eq:diagnosis_start} has terms with $p,q\in k$, terms
-with $p,q\notin k$, and terms with only one of $p$ or $q\in k$.  The
-permutation operator $\Pi_{pq}$ acts trivially on $X_k\ket\psi$ when
-$p,q\in k$ or $p,q\notin k$, so
-\begin{align}
-  \sum_{\substack{k\in\C_N\p{M}\\\p{p,q}\in\C_N\p{2}\\
-      p,q\in k~\t{or}~p,q\notin k}}
+The sum in \eqref{eq:diagnosis_start} has terms with $p,q\notin k$,
+terms with $p,q\in k$, and terms with only one of $p$ or $q\in k$.
+The permutation operator $\Pi_{pq}$ acts trivially on $X_k\ket\psi$
+when $p,q\notin k$, so
+\begin{align}
+  \sum_{\substack{k\in\D_N\p{M}\\\set{p,q}\in\C_N\p{2}\\p,q\notin k}}
   h_{pq} w_k \Pi_{pq} X_k \ket\psi
-  = \sum_{k\in\C_N\p{M}}
-  \sum_{\substack{\p{p,q}\in\C_N\p{2}\\p,q\in k~\t{or}~p,q\notin k}}
-  h_{pq} w_k X_k \ket\psi.
-\end{align}
-For any fixed $k\in\C_N\p{M}$, we can simplify
-\begin{align}
-  \sum_{\substack{\p{p,q}\in\C_N\p{2}\\p,q\in k~\t{or}~p,q\notin k}} h_{pq}
-  = \sum_{\p{p,q}\in\C_N\p{2}} h_{pq}
-  - \sum_{\substack{p\in k\\q\in\ZZ_N\setminus k}} h_{pq}
-  = E_0 - \tilde h_k,
-  &&
-  \tilde h_k \equiv \sum_{\substack{p\in k\\q\in\ZZ_N\setminus k}} h_{pq},
-\end{align}
-where we split a sum over $\p{p,q}\in\C_N\p{2}$ with the restriction
-that $p,q\in k$ or $p,q\notin k$ into an unrestricted sum minus a
-remainder with the complementary restriction; and $E_0$ is the
-interaction energy of PS states $\ket\psi\in\M_0$.  The terms in
-\eqref{eq:diagnosis_start} with $p,q\in k$ or $p,q\notin k$ are then
-\begin{align}
-  \sum_{k\in\C_N\p{M}} \p{E_0 - \tilde h_k} w_k X_k \ket\psi
-  = E_0 X_{\m w} \ket\psi
-  - \sum_{k\in\C_N\p{M}} \p{\tilde{\m h} * \m w}_k X_k \ket\psi,
-\end{align}
-where $\tilde{\m h}$ is a tensor with entries $\tilde h_k$; and
-$\tilde{\m h} * \m w$ denotes an element-wise (Kronecker) product of
-$\tilde{\m h}$ and $\m w$, with entries $\tilde h_k w_k$.  The
-remaining terms in \eqref{eq:diagnosis_start} with only one of
+  = E_0 \tilde X_{\m w} \ket\psi
+  - \sum_{k\in\D_N\p{M}} h^{\t{D}}_k w_k X_k \ket\psi,
+\end{align}
+where $E_0 = \sum_{\set{p,q}\in\C_N\p{2}} h_{pq}$ is the interaction
+energy of PS states $\ket\psi\in\M_0$, and
+\begin{align}
+  h^{\t{D}}_k \equiv
+  \sum_{\substack{\set{p,q}\in\C_N\p{2}\\p\in k~\t{or}~q\in k}} h_{pq}
+  = \sum_{p\in k} h_p - \sum_{\set{a,b}\in\C_M\p{2}} h_{k_ak_b},
+  &&
+  h_p \equiv \sum_{q\in\ZZ_N} h_{pq}.
+\end{align}
+Defining a linear operator $\m h_{\t{D}}$ on the space of
+dimension-$M$ tensors, whose action on $\m w$ is a new dimension-$M$
+tensor with entries
+\begin{align}
+  \m h_{\t{D}}\p{\m w}_k \equiv h^{\t{D}}_k w_k,
+  \label{eq:multi_body_op_diag}
+\end{align}
+we can write the terms in \eqref{eq:diagnosis_start} with
+$p,q\notin k$ as
+\begin{align}
+  \sum_{\substack{k\in\D_N\p{M}\\\set{p,q}\in\C_N\p{2}\\p,q\notin k}}
+  h_{pq} w_k \Pi_{pq} X_k \ket\psi
+  = E_0 \tilde X_{\m w} \ket\psi
+  - \sum_{k\in\D_N\p{M}} \m h_{\t{D}}\p{\m w}_k X_k \ket\psi.
+\end{align}
+The terms in \eqref{eq:diagnosis_start} with $p,q\in k$ are
+\begin{align}
+  \sum_{\substack{k\in\D_N\p{M}\\\set{p,q}\in\C_N\p{2}\\p,q\in k}}
+  h_{pq} w_k \Pi_{pq} X_k \ket\psi
+  = \sum_{k\in\D_N\p{M}} \sum_{\substack{\set{a,b}\in\C_M\p{2}}}
+  h_{k_ak_b} w_{k_{a\lra b}} X_k \ket\psi,
+\end{align}
+where $k_{a\lra b}$ denotes a list that is equal to
+$k=\p{k_1,k_2,\cdots,k_M}$ but with $k_a$ and $k_b$ swapped, and we
+have relabeled $k_a$ and $k_b$ to change $w_k X_{k_{a\lra b}}$ to
+$w_{k_{a\lra b}} X_k$.  To write down these terms succinctly, we
+define a linear operator $\m h_{\t{P}}$ on the space of dimension-$M$
+tensors, whose action on $\m w$ is a new dimension-$M$ tensor with
+entries
+\begin{align}
+  \m h_{\t{P}}\p{\m w}_k
+  \equiv \sum_{\substack{\set{a,b}\in\C_M\p{2}}}
+  h_{k_ak_b} w_{k_{a\lra b}},
+  \label{eq:multi_body_op_perm}
+\end{align}
+such that
+\begin{align}
+  \sum_{\substack{k\in\D_N\p{M}\\\set{p,q}\in\C_N\p{2}\\p,q\in k}}
+  h_{pq} w_k \Pi_{pq} X_k \ket\psi
+  = \sum_{k\in\D_N\p{M}} \m h_{\t{P}}\p{\m w}_k X_k \ket\psi.
+\end{align}
+Finally, the terms in \eqref{eq:diagnosis_start} with only one of
 $p\in k$ or $q\in k$ are
 \begin{align}
-  \sum_{k\in\C_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}} \sum_{q\in k}
-  h_{pq} w_k \Pi_{pq} X_k \ket\psi
-  = \sum_{k\in\C_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}}
-  \sum_{a\in\ZZ_M} h_{p k_a} w_k \Pi_{p k_a} X_k \ket\psi.
-\end{align}
-where $k=\p{k_1,k_2,\cdots,k_M}$, with $k_a$ the $a$-th entry in $k$.
-Defining $k_{a:p}$ to be a list that is equal to $k$ but with the
-$a$-th number $k_a$ replaced by $p$,
-$k_{a:p}\equiv\p{k_1,k_2,\cdots,k_{a-1},p,k_{a+1},\cdots,k_M}$, we can
-simplify
-\begin{align}
-  \Pi_{p k_a} X_k \ket\psi
-  = \Pi_{p k_a} X_k \Pi_{p k_a} \ket\psi
-  = X_{k_{a:p}}\ket\psi,
-\end{align}
-and switch the labels for $p$ and $k_a$ to find that
-\begin{align}
-  \sum_{k\in\C_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}}
+  \sum_{k\in\D_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}}
   \sum_{q\in k} h_{pq} w_k  \Pi_{pq} X_k \ket\psi
-  = \sum_{k\in\C_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}}
-  \sum_{a\in\ZZ_M} h_{pk_a} w_{k_{a:p}} X_k \ket\psi.
-\end{align}
-For each fixed $a$, the sum over $p$ above is essentially a
-contraction of the matrix $\m h$ with the $a$-th index of the tensor
-$\m w$.  Denoting such a contraction by $\m h\c_a\m w$, we can
-therefore write this result as
-\begin{align}
-  \sum_{k\in\C_N\p{M}} \sum_{a\in\ZZ_M}
-  \p{\m h \circ_a \m w}_k X_k \ket\psi,
-\end{align}
-where, for clarity, the matrix elements of $\m h\circ\m w$ are
-\begin{align}
-  \p{\m h \circ_a \m w}_k
+  = \sum_{k\in\D_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}}
+  \sum_{a\in\ZZ_M} h_{pk_a} w_{k_{a:p}} X_k \ket\psi,
+\end{align}
+where $k_{a:p}$ denotes a list that is equal to $k$ but with $k_a$
+replaced by $p$, and we have relabeled $k_a$ and $p$ to change
+$w_k X_{k_{a:p}}$ to $w_{k_{a:p}} X_k$.  For each fixed $a$, the sum
+over $p$ above is essentially a contraction of the matrix $\m h$ with
+the $a$-th index of the tensor $\m w$.  Defining a linear operator
+$\m h_a$ on the space of dimension-$M$ tensors, whose action on $\m w$
+is a new dimension-$M$ tensor with entries
+\begin{align}
+  \m h_a\p{\m w}_k
   \equiv \sum_{\substack{p\in\ZZ_N\\p\notin k}} h_{pk_a} w_{k_{a:p}}
   = \sum_{\substack{p\in\ZZ_N\\p\notin k}}
-  h_{pk_a} w_{k_1 k_2 \cdots k_{a-1} p k_{a+1} \cdots k_M}.
-\end{align}
-Altogether, we thus find that
+  h_{pk_a} w_{k_1 k_2 \cdots k_{a-1} p k_{a+1} \cdots k_M},
+  \label{eq:multi_body_op_contract}
+\end{align}
+we can expand
+\begin{align}
+  \sum_{k\in\D_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}}
+  \sum_{q\in k} h_{pq} w_k  \Pi_{pq} X_k \ket\psi
+  = \sum_{k\in\D_N\p{M}} \sum_{a\in\ZZ_M}
+  \m h_a\p{\m w}_k X_k \ket\psi.
+\end{align}
+Altogether, we find that
 \begin{align}
   H_0 X_{\m w} \ket\psi
-  = E_0 X_{\m w} \ket\psi + \sum_{k\in\C_N\p{M}}
-  \hat{\m h}\p{\m w}_k X_k \ket\psi,
-  &&
+  = E_0 X_{\m w} \ket\psi + X_{\hat{\m h}\p{\m w}} \ket\psi,
+\end{align}
+where
+\begin{align}
   \hat{\m h}\p{\m w}
-  \equiv \sum_{a\in\ZZ_M} \m h \circ_a \m w - \tilde{\m h} * \m w.
-  \label{eq:diagnosis_end}
-\end{align}
-The action of the $M$-body perturbation $X_{\m w}$ on a
-permutationally symmetric state therefore generates an eigenstate of
-$H_0$ with energy $E_0+\Delta$ if the coefficient tensor $\m w$
-satisfies the generalized eigenvalue equation
-$\hat{\m h}\p{\m w} = \Delta \m w$, which can be written in the form
+  \equiv \sum_{a\in\ZZ_M} \m h_a\p{\m w}
+  + \m h_{\t{P}}\p{\m w} - \m h_{\t{D}}\p{\m w},
+\end{align}
+is defined in terms of the tensors $\m h_a\p{\m w}$,
+$\m h_{\t{P}}\p{\m w}$, and $\m h_{\t{D}}\p{\m w}$, whose entries are
+respectively provided in \eqref{eq:multi_body_op_contract},
+\eqref{eq:multi_body_op_perm}, and \eqref{eq:multi_body_op_diag}.  The
+action of the $M$-body perturbation $X_{\m w}$ on a permutationally
+symmetric state therefore generates an eigenstate of $H_0$ with energy
+$E_0+\Delta$ if the coefficient tensor $\m w$ satisfies the
+generalized eigenvalue equation $\hat{\m h}\p{\m w} = \Delta \m w$,
+which can be written in the form
 \begin{align}
   \hat{\m h} \c \v{\m w} = \Delta \v{\m w},
+  &&
+  \v{\m w} \equiv \sum_{k\in\C_N\p{M}} w_k \ket{k},
   \label{eq:multi_body_eig_mat}
@@
 \begin{align}
-  \hat{\m h} \equiv \sum_{k\in\C_N\p{M}}
-  \sum_{\substack{p\in\ZZ_N\\p\notin k}}
-  \sum_{a\in\ZZ_M} h_{pk_a} \op{k}{k_{a:p}}
-  - \sum_{k\in\C_N\p{M}} \tilde h_k \op{k},
-  &&
-  \v{\m w} \equiv \sum_{k\in\C_N\p{M}} w_k \ket{k}.
-\end{align}
-If $M=1$, then
-$\p{\m h\circ\m w}_k=\sum_{\ell\in\ZZ_N} h_{k\ell} w_\ell$ and
-$\tilde h_k = h_k \equiv \sum_{\ell\in\ZZ_N} h_{k\ell}$, so
-\begin{align}
-  H_0 X_{\m w} \ket\psi
-  = E_0 X_{\m w} \ket\psi + \sum_{k\in\ZZ_N}
-  \p{\sum_{\ell\in\ZZ_N} h_{k\ell} w_\ell - h_k w_k} X_k \ket\psi,
-\end{align}
-which is precisely what we found in
-\eqref{eq:single_body_diagnosis_end}.  If $M=2$, then
-$\p{\m h\circ\m w}_{pq} = \sum_{k\in\ZZ_N}
-\p{h_{pk}w_{kq}+h_{qk}w_{pk}}$, so
-\begin{align}
-  H_0 X_{\m w} \ket\psi
-  = E_0 X_{\m w} \ket\psi + \sum_{\p{p,q}\in\C_N\p{2}}
-  \sp{\sum_{k\in\ZZ_N}\p{h_{pk} w_{kq} + h_{qk} w_{pk}}
-    - \tilde h_{pq} w_{pq}}
-  X_k \ket\psi,
-\end{align}
-where $\tilde h_{pq}=h_p+h_q-2h_{pq}$.
+  \hat{\m h} \equiv \sum_{k\in\D_N\p{M}} \ket{k}
+  \sp{\sum_{a\in\ZZ_M} \sum_{\substack{p\in\ZZ_N\\p\notin k}}
+    h_{pk_a} \bra{k_{a:p}}
+    + \sum_{\set{a,b}\in\C_M\p{2}} h_{k_ak_b} \bra{k_{a\lra b}}
+    - h^{\t{D}}_k \bra{k}}.
+\end{align}
+If $M=1$, then $h^{\t{D}}_k=h_k$, so
+\begin{align}
+  \hat{\m h} \stackrel{M=1}{=} \sum_{k\in\ZZ_N} \ket{k}
+  \sp{\sum_{\substack{p\in\ZZ_N\\p\ne k}} h_{pk} \bra{p} - h_k \bra{k}}
+  = \m h - \diag\v h,
+\end{align}
+which is precisely what we found in \eqref{eq:single_body_eig}.  If
+$M=2$, then $h_{pq}^{\t{D}}=h_p+h_q-h_{pq}$, so
+\begin{align}
+  \hat{\m h} \stackrel{M=2}{=}
+  \sum_{\p{k,\ell}\in\D_N\p{2}} \ket{k,\ell}
+  \sp{\sum_{\substack{p\in\ZZ_N\\p\notin\set{k,\ell}}}
+    \p{ h_{pk} \bra{p,\ell} + h_{p\ell} \bra{k,p} }
+    + h_{k\ell} \bra{\ell,k}
+    - \p{h_k + h_\ell - h_{k\ell}} \bra{k,\ell}}.
+\end{align}
 
@@
 \label{sec:symmetries}
+
+\red{[TODO: fix remainder of this section to account for perturbations
+  that do not obey permutational symmetry]}
 
@@
 matrix $\hat{\m h}$ with dimensions
-${N\choose M}\times{N\choose M}\sim N^M\times N^M$.  If the
-interaction Hamiltonian $H_0$ has any spatial symmetries, however,
-then we can use these symmetries to reduce the complexity of this
-eigenvalue problem.  Translational invariance, for example, can reduce
-this problem to that of diagonalizing a matrix with dimensions
+$N!_M\times N!_M\sim N^M\times N^M$, where
+$N!_M\equiv\prod_{j=0}^{M-1}\p{N-j}$ denotes a falling factorial.
+
+\vspace{3cm}
+
+If the interaction Hamiltonian $H_0$ has any spatial symmetries,
+however, then we can use these symmetries to reduce the complexity of
+this eigenvalue problem.  Translational invariance, for example, can
+reduce this problem to that of diagonalizing a matrix with dimensions
 $\le{N-1\choose M-1}\times{N-1\choose M-1}\sim N^{M-1}\times N^{M-1}$,
@@
 equivalence class $s\in\S_2$ would consist of all pairs
-$\p{p,q}\in\C_N\p{2}$ that are a fixed distance $\abs{p-q}=d_s$ apart,
-i.e.~$s=\set{\p{p,q}\in\C_N\p{2}:\abs{p-q}=d_s}$.
+$\set{p,q}\in\C_N\p{2}$ that are a fixed distance $\abs{p-q}=d_s$
+apart, i.e.~$s=\set{\set{p,q}\in\C_N\p{2}:\abs{p-q}=d_s}$.
 
@@
 \begin{multline}
-  X_{\m v} Y_{\m w}
-  = \sp{\sum_{\p{k,\ell},\p{p,q}\in\D_N\p{2}}
+  X_{\m v} Y_{\m w} = \sp{\sum_{\set{k,\ell},\set{p,q}\in\C_N\p{2}}
     v_{k\ell} w_{pq} X_{k\ell} Y_{pq}} \\
@@
 \begin{align}
-  \v{\m m} \equiv \sum_{\p{p,q}\in\C_N\p{2}} m_{pq} \ket{p,q},
+  \v{\m m} \equiv \sum_{\set{p,q}\in\C_N\p{2}} m_{pq} \ket{p,q},
   &&
@@
   &&
-  \col{m} \equiv \sum_{\p{p,q}\in\C_N\p{2}} m_{pq},
+  \col{m} \equiv \sum_{\set{p,q}\in\C_N\p{2}} m_{pq},
 \end{align}
@@
 \begin{multline}
-  \sum_{\p{k,\ell},\p{p,q}\in\C_N\p{2}} v_{k\ell} w_{pq}
+  \sum_{\p{k,\ell},\set{p,q}\in\C_N\p{2}} v_{k\ell} w_{pq}
   Z_k Z_\ell Z_p Z_q \\
@@
   Z_{\m u\m v\m w}
-  \equiv \sum_{\p{k,\ell},\p{p,q},\p{r,s}\in\C_N\p{2}}
+  \equiv \sum_{\set{k,\ell},\set{p,q},\set{r,s}\in\C_N\p{2}}
     u_{k\ell} v_{pq} w_{rs} Z_k Z_\ell Z_p Z_q Z_r Z_s,
@@
 \begin{align}
-  \v{\m m} \equiv \sum_{\p{p,q}\in\C_N\p{2}} m_{pq} \ket{\set{p,q}},
+  \v{\m m} \equiv \sum_{\set{p,q}\in\C_N\p{2}} m_{pq} \ket{\set{p,q}},
   &&
@@
     + v_p w_{pq} u_{pq} + w_p u_{pq} v_{pq}}
-  + 4 \sum_{\p{p,q}\in\C_N\p{2}} u_{pq} v_{pq} w_{pq}.
+  + 4 \sum_{\set{p,q}\in\C_N\p{2}} u_{pq} v_{pq} w_{pq}.
 \end{multline}

### 2020-03-26 collective_sun/sun_perturbation.tex
finished implementing fixes for multi-body operators without permutational symmetry

@@
 \begin{align}
-  H_0 = \sum_{\substack{p,q\in\ZZ_N\\p<q}} h_{pq} \Pi_{pq},
+  H_0 = \sum_{\set{p,q}\in\C_N\p{2}} h_{pq} \Pi_{pq},
   &&
@@
 \end{align}
-where $p,q$ index individual spins; $\ZZ_k$ is the set of integers
-modulo $k$; $h_{pq}$ are scalar coupling constants; $\mu,\nu$ index
-states in an orthonormal basis for the $n$-dimensional Hilbert space
-of a single spin; the operator $s_{\mu\nu}^{(p)}\equiv\op{\mu}{\nu}_p$
-flips the state of spin $p$ to $\ket\nu$ from $\ket\mu$; and the
-operator $\Pi_{pq}$ permutes the states of spins $p$ and $q$.  The
-permutation operator $\Pi_{pq}$ is essentially a multilevel
-generalization of the SU(2)-symmetric spin interaction
-$\v s_p\c\v s_q$, with $\v s\equiv\p{\v s_\x,s_\y,s_\z}$,
+where $p,q$ index individual spins; $\C_N\p{2}$ is the set of all
+subsets (``choices'') of two spins out of $N$; $h_{pq}$ are scalar
+coupling constants; $\mu,\nu$ index states in an orthonormal basis for
+the $n$-dimensional Hilbert space of a single spin; $\ZZ_n$ is the set
+of integers modulo $n$; the operator
+$s_{\mu\nu}^{(p)}\equiv\op{\mu}{\nu}_p$ flips the state of spin $p$ to
+$\ket\nu$ from $\ket\mu$; and the operator $\Pi_{pq}$ permutes the
+states of spins $p$ and $q$.  The permutation operator $\Pi_{pq}$ is
+essentially a multilevel generalization of the SU(2)-symmetric spin
+interaction $\v s_p\c\v s_q$, with $\v s\equiv\p{\v s_\x,s_\y,s_\z}$,
 $s_\alpha\equiv\sigma_\alpha/2$, and $\sigma_\alpha$ a single-spin
@@
 $\Pi_{pq}$.  We will assume that all $h_{pq}<0$ throughout these
-notes, but comment that this restriction can be relaxed to the
-assumption that the initial state of all spins is permutationally
-symmetric (e.g.~a spin-polarized state).  We will also assume that the
-PS manifold $\M_0$ is gapped away from all orthogonal states by an
+notes, but note that this restriction can be relaxed to the assumption
+that the initial state of all spins is permutationally symmetric
+(e.g.~a spin-polarized state).  We will also assume that the PS
+manifold $\M_0$ is gapped away from all orthogonal states by an
 interaction energy $\Delta_{\t{gap}}>0$.  Power-law couplings of the
-form $h_{pq}\sim-1/\abs{r_p-r_q}^\alpha$ with positions $r_p,r_q$ on a
-$D$-dimensional lattice, for example, always yield a non-vanishing
-spectral gap $\Delta_{\t{gap}}$ for long-range interactions with
-$\alpha\le D$ (see Appendix \ref{sec:spectral_gap}).
+form $h_{pq}\sim-1/\abs{p-q}^\alpha$ on a $D$-dimensional lattice, for
+example, always yield a non-vanishing spectral gap $\Delta_{\t{gap}}$
+for long-range interactions with $\alpha\le D$ (see Appendix
+\ref{sec:spectral_gap}).
 
@@
 \begin{align}
-  E_0 = \sum_{\substack{p,q\in\ZZ_N\\p<q}} h_{pq}.
+  E_0 = \sum_{\set{p,q}\in\C_N\p{2}} h_{pq}.
 \end{align}
@@
 \end{align}
-is the set of subsets (``choices'') of $M$ elements from $\ZZ_N$.  We
-can then expand
+is the set of all subsets (``choices'') of $M$ elements from $\ZZ_N$.
+We can then expand
 \begin{align}
@@
 \end{align}
-For simplicity, we will assume throughout this work that $\hat\O$ is
-homogeneous in the sense that $M$ is the same for all
-$\p{X,\m w}\in\O$, with the understanding that a generalization to
-inhomogeneous operators $\hat\O$ is straightforward.
+To simplify language and notation, we will assume throughout this work
+that $\hat\O$ is homogeneous in the sense that $M$ is the same for all
+$\p{X,\m w}\in\O$.  A generalization to the case of inhomogeneous
+operators $\hat\O$ is straightforward.
 
@@
 \end{align}
-with $\abs{\D_N\p{M}}=M!\times{N\choose M}=\prod_{j=0}^{M-1}\p{N-j}$
-and $\abs{\pi\p{M}}=M!$.  An immediate consequence of
-\eqref{eq:H_eff_1} is the fact that perturbing SU(2)-symmetric
-interactions by two-body $s_\z\otimes s_\z$ interactions yields the
-one-axis twisting (OAT) Hamiltonian\cite{kitagawa1993squeezed,
-  ma2011quantum} $H_{\t{OAT}}=\chi S_\z^2$ at first order in
-perturbation theory, with the OAT strength $\chi$ equal to the mean
-coefficient of the two-body $s_\z\otimes s_\z$ terms.
-
-\red{[TODO: fix remainder of this section to account for perturbations
-  that do not obey permutational symmetry]}
+with $\abs{\D_N\p{M}}=\prod_{j=0}^{M-1}\p{N-j}$ and
+$\abs{\pi\p{M}}=M!$.  An immediate consequence of \eqref{eq:H_eff_1}
+is the fact that perturbing SU(2)-symmetric interactions by two-body
+$s_\z\otimes s_\z$ interactions yields the one-axis twisting (OAT)
+Hamiltonian\cite{kitagawa1993squeezed, ma2011quantum}
+$H_{\t{OAT}}=\chi S_\z^2$ at first order in perturbation theory, with
+the OAT strength $\chi$ equal to the mean coefficient of the two-body
+$s_\z\otimes s_\z$ terms.
 
@@
 $\P_\Delta$ onto the manifold $\E_\Delta$ of states with excitation
-energy $\Delta$ (with respect to $H_0$).  For each term $X_{\m w}$ of
-$\hat\O$, this projector essentially picks off the part of $X_{\m w}$
-that is strictly off-diagonal with respect to the ground- and
-excited-state manifolds $\E_0$ and $\E_\Delta$.  To decompose
-$X_{\m w}$ into components that generate states of definite excitation
-energy when acting on PS states $\M_0$, we pick an arbitrary state
-$\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:eigenstates})
+energy $\Delta$.  For each term $X_{\m w}$ of $\hat\O$, this projector
+essentially picks off the part of $X_{\m w}$ that is strictly
+off-diagonal with respect to the ground- and excited-state manifolds
+$\E_0$ and $\E_\Delta$.  To decompose $X_{\m w}$ into components that
+generate states of definite excitation energy when acting on PS states
+$\M_0$, we pick an arbitrary state $\ket\psi\in\M_0$ and expand (see
+Appendix \ref{sec:eigenstates})
 \begin{align}
@@
 where $\hat{\m h}$ is a linear map on the space of dimension-$M$
-tensors, determined entirely by the couplings $h_{pq}$ of the
-interaction Hamiltonian $H_0$.  The image $\hat{\m h}\p{\m w}$ of
-$\m w$ under $\hat{\m h}$ is then itself a dimension-$M$ tensor that
-defines the $M$-body operator $X_{\hat{\m h}\p{\m w}}$.  To write out
-$\hat{\m h}\p{\m w}$ explicitly, we first define a matrix
+tensors, determined entirely by the couplings $h_{pq}$ in $H_0$.  The
+image $\hat{\m h}\p{\m w}$ of $\m w$ under $\hat{\m h}$ is then itself
+a dimension-$M$ tensor that defines the $M$-body operator
+$X_{\hat{\m h}\p{\m w}}$.  To write out $\hat{\m h}\p{\m w}$
+explicitly, we first define a matrix
 $\m h\equiv\sum_{p,q\in\ZZ_N} h_{pq} \op{p}{q}$ of all couplings in
@@
 contraction of $\m h$ with the $a$-th index of $\m w$ by
-$\m h \circ_a\m w$, such that the $k$-th entry of $\m h \circ_a\m w$
-is
-\begin{align}
-  \p{\m h \circ_a \m w}_k
-  = \sum_{p\in\ZZ_N\setminus k} h_{k_a p} w_{k_{a:p}},
-\end{align}
-where $k\equiv\p{k_1,k_2,\cdots,k_M}$ is a list of indices;
-$k_{a:p}\equiv\p{k_1,k_2,\cdots,k_{a-1},p,k_{a+1},\cdots,k_M}$ is the
-same list $k$, but with the $a$-th index $k_a$ replaced by $p$; and we
-explicitly exclude terms with $p\in k$ that vanish due to the fact
-that $\m h$ and $\m w$ are strictly off-diagonal.  We then define a
-dimension-$M$ tensor $\tilde{\m h}$ with entries
-\begin{align}
-  \tilde h_k \equiv \sum_{\substack{p\in k\\q\in\ZZ_N\setminus k}} h_{pq}
-\end{align}
-for all $k\in\ZZ_N^M$, and we denote an element-wise (Kronecker)
-product of $\tilde{\m h}$ and $\m w$ by $\tilde{\m h} * \m w$, with
-entries $\tilde h_k w_k$.  These definitions allow us to expand
+$\m h_a\p{\m w}$, such that the $k$-th entry of $\m h_a\p{\m w}$ is
+\begin{align}
+  \m h_a\p{\m w}_k
+  \equiv \sum_{\substack{p\in\ZZ_N\\p\notin k}}
+  h_{pk_a} w_{k_1 k_2 \cdots k_{a-1} p k_{a+1} \cdots k_M}
+  = \sum_{\substack{p\in\ZZ_N\\p\notin k}} h_{pk_a} w_{k_{a:p}},
+\end{align}
+where if $k=\p{k_1,k_2,\cdots,k_M}$, then $k_{a:p}$ is the same list
+but with the $a$-th entry $k_a$ replaced by $p$.  We similarly define
+a tensor $\m h_{\t{P}}\p{\m w}$ with entries
+\begin{align}
+  \m h_{\t{P}}\p{\m w}_k
+  \equiv \sum_{\set{a,b}\in\C_M\p{2}} h_{k_ak_b} w_{k_{a\lra b}},
+\end{align}
+where $k_{a\lra b}$ is the same as $k$ but with the placement of $k_a$
+and $k_b$ swapped.  Finally, we define a tensor $\m h_{\t{D}}\p{\m w}$
+with entries
+\begin{align}
+  \m h_{\t{D}}\p{\m w}_k \equiv h^{\t{D}}_k w_k,
+  &&
+  h^{\t{D}}_k \equiv \sum_{p\in k} h_p
+  - \sum_{\set{a,b}\in\C_M\p{2}} h_{k_ak_b},
+  &&
+  h_p \equiv \sum_{q\in\ZZ_N} h_{pq}.
+\end{align}
+These definitions allow us to expand
 \begin{align}
   \hat{\m h}\p{\m w}
-  \equiv \sum_{a\in\ZZ_M} \m h \circ_a \m w - \tilde{\m h} * \m w.
-  \label{eq:multi_body_eig_op}
+  \equiv \sum_{a\in\ZZ_M} \m h_a\p{\m w}
+  + \m h_{\t{P}}\p{\m w} - \m h_{\t{D}}\p{\m w}.
 \end{align}
@@
 and eigenvectors of $\hat{\m h}$ nominally requires diagonalizing a
-matrix of dimensions
-${N\choose M}\times{N\choose M}\sim N^M\times N^M$, but spatial
-symmetries of $H_0$ such as translational invariance or isotropy can
-drastically reduce the complexity of this eigenvalue problem.  We
-discuss this reduction in Appendix \ref{sec:symmetries}.
+matrix of dimensions $\abs{\D_N\p{M}}\times\abs{\D_N\p{M}}$, but
+spatial symmetries of $H_0$ and $X_{\m w}$ such as translational
+invariance or isotropy can drastically reduce the complexity of this
+eigenvalue problem.  We discuss these reductions in Appendix
+\ref{sec:symmetries}.
 
@@
 which implies that the action of $X_{\m w_\Delta}$ on a PS state
-$\ket\psi\in\M_0$ generates a state of definite interaction energy
+$\ket\psi\in\M_0$ generates a state with definite interaction energy
 $E_0+\Delta$.  It follows that
@@
 \end{align}
-The product $\P_0 X_{\m v_\Delta} Y_{\m w_\Delta} \P_0$ generally
-includes $r$-body operators for all positive integers $r\le2M$.  We
-provide a general prescription for simplifying such products in
-Appendix \ref{sec:operator_product}, and explicitly work out the
-single-body ($M=1$) and two-body ($M=2$) cases in Appendices
+Simplifying product $\P_0 X_{\m v_\Delta} Y_{\m w_\Delta} \P_0$
+generally yields $r$-body operators for all positive integers
+$r\le2M$.  We provide a general prescription for simplifying such
+products in Appendix \ref{sec:operator_product}, and explicitly work
+out single-body ($M=1$) and two-body ($M=2$) cases in Appendices
 \ref{sec:single_pair_prod} and \ref{sec:two_pair_prod}.  If $X$ is a
@@
 $X_{\m w}$ onto $\ul{\M}_{r_{\t{max}}}$ for some $r_{\t{max}}\ge M$,
-resulting in a theory analogous to first-order perturbation theory on
-$\ul{\M}_{r_{\t{max}}}$.  In the perturbative regime
+resulting in a theory that resembles first-order perturbation theory
+on $\ul{\M}_{r_{\t{max}}}$.  In the perturbative regime
 $\norm{X_{\m w}}<\Delta_{\t{gap}}/2$, such a theory captures all
@@
 \end{align}
-where the domain of $d$ is defined in terms of integers modulo $L$,
-i.e.~$\ZZ_L\simeq\set{0,1,\cdots,L-1}$ with the relation $\simeq$
-denoting an association that ignores the cyclic structure of $\ZZ_L$;
+where $\ZZ_L$ denotes the integers modulo $L$;
 $k_{\t{SE}}\in\ZZ_L^D\times2\pi/L$ is a wavenumber for a
@@
 this behavior, we note that the asymptotic dependence of the integral
-$I_{D\epsilon}$ on $\epsilon$ is determined by the behavior of its
+$\I_{D\epsilon}$ on $\epsilon$ is determined by the behavior of its
 integrand when $\abs{x}\sim\epsilon$, in which case
@@
 In this section, we simplify the products of the form
-$H_0X_{\m w}\ket\psi$ to arrive at the expansion in
-\eqref{eq:diagnosis}, which defines conditions under which a
-multi-body perturbation $X_{\m w_\Delta}$ generates eigenstates of the
-SU($n$)-symmetric interaction Hamiltonian $H_0$ in \eqref{eq:H_0} when
-acting on a permutationally symmetric state $\ket\psi\in\M_0$.  Here
-$X_{\m w}$ is an $M$-body perturbation defined by
+$H_0X_{\m w}\ket\psi$ to find conditions under which a multi-body
+perturbation $X_{\m w}$ generates eigenstates of the SU($n$)-symmetric
+interaction Hamiltonian $H_0$ in \eqref{eq:H_0} when acting on a
+permutationally symmetric state $\ket\psi\in\M_0$.  Here $X_{\m w}$ is
+an $M$-body perturbation defined by
 \begin{align}
   X_{\m w}
-  \equiv \f1{M!} \sum_{k\in\D_N\p{M}} w_k X_k
-  = \f1{M!} \sum_{\sigma\in\pi\p{M}} \sum_{k\in\C_N\p{M}}
-  w_{\sigma\p{k}} X_{\sigma\p{k}},
+  \equiv \f1{M!} \sum_{k\in\D_N\p{M}} w_k X_k,
 \end{align}
@@
 a list of $M$ spins that $X_k$ acts on; $\m w$ is a dimension-$M$
-(i.e.~$M$-index) tensor with scalar entries $w_k$; $\D_N\p{M}$ is the
-strictly off-diagonal part of $\ZZ_N^M$, defined in
-\eqref{eq:off_diags}; $\pi\p{M}$ is the permutation group of $\ZZ_M$;
-and $\C_N\p{M}$ is the set of subsets (``choices'') of $M$ elements
-from $\ZZ_N$, defined in \eqref{eq:choices}.  For shorthand, we will
-define $\tilde X_{\m w}\equiv M! X_{\m w}$.
+(i.e.~$M$-index) tensor with scalar entries $w_k$; and $\D_N\p{M}$ is
+the strictly off-diagonal part of $\ZZ_N^M$, defined in
+\eqref{eq:off_diags}.
 
@@
 \begin{align}
-  H_0 \tilde X_{\m w} \ket\psi
+  H_0 X_{\m w} \ket\psi
   = \sum_{\substack{k\in\ZZ_N\\\set{p,q}\in\C_N\p{2}}}
@@
 $k\notin\set{p,q}$, the permutation operator $\Pi_{pq}$ commutes with
-$X_k$ and annihilates on $\ket\psi$, so
+$X_k$ and annihilates on $\ket\psi$, allowing us to simplify
 \begin{align}
   \sum_{k\in\ZZ_N} \sum_{\substack{\set{p,q}\in\C_N\p{2}\\k\notin\set{p,q}}}
-  h_{pq} w_k X_k \ket\psi
-  = E_0 \tilde X_{\m w} \ket\psi
+  h_{pq} w_k \Pi_{pq} X_k \ket\psi
+  = E_0 X_{\m w} \ket\psi
   - \sum_{k\in\ZZ_N} \sum_{\substack{\set{p,q}\in\C_N\p{2}\\k\in\set{p,q}}}
@@
 where $E_0 = \sum_{\set{p,q}\in\C_N\p{2}} h_{pq}$ is the interaction
-energy of PS states $\ket\psi\in\M_0$.  Simplifying
+energy of PS states $\ket\psi\in\M_0$, and
 \begin{align}
@@
 \end{align}
-we can write
+which implies
 \begin{align}
@@
   h_{pq} w_k X_k \ket\psi
-  = E_0 \tilde X_{\m w} \ket\psi
+  = E_0 X_{\m w} \ket\psi
   - \sum_{k\in\ZZ_N} h_k w_k X_k \ket\psi.
@@
 \end{align}
-where the imaginary terms vanish because $h_{0,d}=h_{0,-d}$.  The
-remainder of \eqref{eq:single_body_eig} that we need to sort out is
-$\diag\v h$, where all
+where the imaginary contributions vanish because $h_{0,d}=h_{0,-d}$.
+The remainder of \eqref{eq:single_body_eig} that we need to sort out
+is $\diag\v h$, where all
 $h_p=\sum_{q\in\ZZ_L^D}h_{pq}=\sum_{d\in\ZZ_L^D}h_{0,d}$ are equal,
@@
 \subsection{Multi-body excitations}
-\label{sec:multi_body_eigenstates}
 
@@
   = \sum_{\substack{k\in\D_N\p{M}\\\set{p,q}\in\C_N\p{2}}}
-  h_{pq} w_k \Pi_{pq} X_k \ket\psi.
+  h_{pq} w_k \Pi_{pq} X_k \ket\psi,
   \label{eq:diagnosis_start}
 \end{align}
-The sum in \eqref{eq:diagnosis_start} has terms with $p,q\notin k$,
-terms with $p,q\in k$, and terms with only one of $p$ or $q\in k$.
-The permutation operator $\Pi_{pq}$ acts trivially on $X_k\ket\psi$
-when $p,q\notin k$, so
+where to avoid factors of $M!$ in this section, we define
+$\tilde X_{\m w}\equiv M! X_{\m w}$.  The sum in
+\eqref{eq:diagnosis_start} has terms with $p,q\notin k$, terms with
+$p,q\in k$, and terms with only one of $p$ or $q\in k$.  The
+permutation operator $\Pi_{pq}$ acts trivially on $X_k\ket\psi$ when
+$p,q\notin k$, allowing us to simplify
 \begin{align}
@@
   h_p \equiv \sum_{q\in\ZZ_N} h_{pq}.
+  \label{eq:multi_body_op_diag_element}
 \end{align}
@@
   + \m h_{\t{P}}\p{\m w} - \m h_{\t{D}}\p{\m w},
+  \label{eq:multi_body_op}
 \end{align}
@@
   &&
-  \v{\m w} \equiv \sum_{k\in\C_N\p{M}} w_k \ket{k},
-  \label{eq:multi_body_eig_mat}
+  \v{\m w} \equiv \sum_{k\in\D_N\p{M}} w_k \ket{k},
+  \label{eq:multi_body_problem}
 \end{align}
@@
     - h^{\t{D}}_k \bra{k}}.
-\end{align}
-If $M=1$, then $h^{\t{D}}_k=h_k$, so
+  \label{eq:multi_body_mat}
+\end{align}
+If the coefficient tensor $\m w$ is permutationally symmetric, meaning
+that $\m w$ is invariant under arbitrary permutations of its tensor
+factors, then this symmetry is preserved by $\hat{\m h}$.  In this
+case, we can replace sums over $k\in\D_N\p{M}$ in
+\eqref{eq:multi_body_problem} and \eqref{eq:multi_body_mat} by sums
+over $k\in\C_N\p{M}$, and replace vectors
+$\ket{\p{k_1,k_2,\cdots,k_M}}\to\ket{\set{k_1,k_2,\cdots,k_M}}$, such
+that e.g.~$\ket{k_{a\lra b}}=\ket{k}$.  These replacements reduce the
+size of $\hat{\m h}$ from $N!_M\times N!_M$ to
+${N\choose M}\times{N\choose M}$, where
+$N!_M=M!\times{N\choose M}=\prod_{j=0}^{M-1}\p{N-j}$ denotes a falling
+factorial.  We discuss further reduction of the eigenvalue problem in
+\eqref{eq:multi_body_problem} by use of additional symmetries in
+Section \ref{sec:symmetries}.
+
+Some special cases: if $M=1$, then $h^{\t{D}}_k=h_k$, so
 \begin{align}
@@
 which is precisely what we found in \eqref{eq:single_body_eig}.  If
-$M=2$, then $h_{pq}^{\t{D}}=h_p+h_q-h_{pq}$, so
+$M=2$, then $h_{k\ell}^{\t{D}}=h_k+h_\ell-h_{k\ell}$, so
 \begin{align}
@@
 
-\red{[TODO: fix remainder of this section to account for perturbations
-  that do not obey permutational symmetry]}
-
-Solving the eigenvalue problem in \eqref{eq:multi_body_eig_mat} to
-find $M$-body perturbations $X_{\m w}$ that generate states of
-definite interaction energy nominally requires diagonalizing the
-matrix $\hat{\m h}$ with dimensions
-$N!_M\times N!_M\sim N^M\times N^M$, where
-$N!_M\equiv\prod_{j=0}^{M-1}\p{N-j}$ denotes a falling factorial.
-
-\vspace{3cm}
-
-If the interaction Hamiltonian $H_0$ has any spatial symmetries,
-however, then we can use these symmetries to reduce the complexity of
-this eigenvalue problem.  Translational invariance, for example, can
-reduce this problem to that of diagonalizing a matrix with dimensions
-$\le{N-1\choose M-1}\times{N-1\choose M-1}\sim N^{M-1}\times N^{M-1}$,
-and isotropy can further reduce the problem size by factors that
-depend on the lattice geometry.  The complexity of diagonalizing a
-matrix with dimensions $d\times d$ is $O\p{d^3}$, so even a reduction
-of $d$ by a factor of 2 yields practically significant gains (a factor
-of $2^3=8$) in the runtime of diagonalization.  Here, we provide a
-general prescription for exploiting the spatial symmetries of $H_0$ to
-simplify the eigenvalue problem in \eqref{eq:multi_body_eig_mat}.
+Solving the eigenvalue problem in \eqref{eq:multi_body_problem} to
+find $M$-body operators $X_{\m w}$ that generate states of definite
+interaction energy nominally requires diagonalizing the matrix
+$\hat{\m h}$ with dimensions $N!_M\times N!_M\sim N^M\times N^M$,
+where $N!_M=\prod_{j=0}^{M-1}\p{N-j}$ denotes a falling factorial.
+Spatial symmetries of the interaction Hamiltonian $H_0$, however,
+generally endow $\hat{\m h}$ with a block-diagonal structure that can
+be used to reduce the complexity of this eigenvalue problem.
+Translational invariance, for example, can reduce this problem to that
+of diagonalizing a matrix with dimensions
+$\sim N^{M-1}\times N^{M-1}$, and isotropy can further reduce the
+problem size by factors that depend on the lattice geometry.  The
+complexity of diagonalizing a matrix with dimensions $d\times d$ is
+$O\p{d^3}$, so even a reduction of $d$ by a factor of 2 yields
+practically significant gains (a factor of $2^3=8$) in the runtime of
+diagonalization.  Here, we provide a general prescription for
+exploiting the spatial symmetries of $H_0$ to simplify the eigenvalue
+problem in \eqref{eq:multi_body_problem}.
 
 Our task is essentially to express the multi-body eigenvalue problem
-in \eqref{eq:multi_body_eig_mat} in a basis that respects the spatial
-symmetries of $H_0$.  These symmetries essentially break the set of
-choices $\C_N\p{M}$ into equivalence classes $\S_M$\footnote{For
-  brevity, we will suppress the explicit dependence of $\S_M$ on $N$,
-  as well as the dependence of $\S_M$ on any other factors such as
-  lattice geometry.}, where elements $s\in\S_M$ are subsets
-$s\subset\C_N\p{M}$ that are invariant under the appropriate symmetry
-group action.  In the case of a translationally invariant and
-isotropic system with two-body perturbations, for example, each
-equivalence class $s\in\S_2$ would consist of all pairs
-$\set{p,q}\in\C_N\p{2}$ that are a fixed distance $\abs{p-q}=d_s$
-apart, i.e.~$s=\set{\set{p,q}\in\C_N\p{2}:\abs{p-q}=d_s}$.
+in \eqref{eq:multi_body_problem} in a basis that respects the spatial
+symmetries of $H_0$.  These symmetries essentially break the set
+$\D_N\p{M}$ into equivalence classes $\S_M$\footnote{For brevity, we
+  will suppress the explicit dependence of $\S_M$ on $N$, as well as
+  the dependence of $\S_M$ on any other factors such as lattice
+  geometry.}, where elements $s\in\S_M$ are subsets
+$s\subset\D_N\p{M}$ that are invariant under the appropriate symmetry
+group action.  In the case of $M=2$ in a translationally invariant and
+isotropic system, for example, each equivalence class $s\in\S_2$ would
+consist of all pairs $\p{p,q}\in\D_N\p{2}$ that are a fixed distance
+$\abs{p-q}=d_s$ apart,
+i.e.~$s=\set{\p{p,q}\in\D_N\p{2}:\abs{p-q}=d_s}$.  If the tensor
+$\m w$ is invariant under arbitrary permutations of its tensor
+factors, this symmetry is preserved by $\hat{\m h}$, and we can place
+all $k\in\D_N\p{M}$ that are equal up to permutation into the same
+equivalence class.
 
@@
 $s\in\S_M$, we then define a uniform superposition $\ket{s}$ over all
-choices $k\in s$, as well as the projector $\P_\S$ onto the space of
-uniform superpositions:
+$k\in s$, as well as a projector $\P_\S$ onto the space of uniform
+superpositions:
 \begin{align}
@@
 \end{align}
-where $\abs{s}$ is the number of elements in $s$.  We then expand the
-coefficient vector $\v{\m w}$ in this symmetrized basis:
-\begin{align}
-  \v{\m w} = \sum_{k\in\C_N\p{M}} w_k \ket{k}
+where $\abs{s}$ is the number of elements in $s$.  In turn, we expand
+the coefficient vector $\v{\m w}$ in this symmetrized basis:
+\begin{align}
+  \v{\m w} = \P_{\S_M} \v{\m w}
+  = \sum_{s\in\S_M} \sum_{k\in s} w_k \ket{s} \bk{s|k}
   = \sum_{s\in\S_M} w_s \sqrt{\abs{s}} \ket{s},
 \end{align}
-where $w_s\equiv w_k$ for any $k\in s$.  Finally, we define the
+where all $w_k$ with $k\in s$ are equal by assumption, so we can
+define $w_s\equiv w_k$ for any $k\in s$.  Finally, we expand the
 projection of $\hat{\m h}$ onto the symmetric subspace:
 \begin{align}
-  \hat{\m h}_{\S_M}
-  \equiv \P_{\S_M} \hat{\m h} \P_{\S_M}
-  = \sum_{s\in\S_M} \ket{s} \sp{
-    \sum_{k\in s} \sum_{\substack{p\in\ZZ_N\\p\notin k}} \sum_{a\in\ZZ_M}
-    \f{h_{p k_a}}{\sqrt{\abs{s}\abs{\sp{k_{a:p}}}}} \bra{\sp{k_{a:p}}}
-    - \tilde h_s \bra{s}},
-\end{align}
-where $\tilde h_s\equiv \tilde h_k$ for any $k\in s$; $k_a$ is the
-$a$-th element of $k$;
-$k_{a:p}\equiv\p{k_1,k_2,\cdots,k_{a-1},p,k_{a+1},\cdots,k_M}$ is $k$,
-but with the $a$-th number $k_a$ replaced by $p$; and $\sp{k_{a:p}}$
-is the equivalence class of $k_{a:p}$.  The reduced multi-body
+  \P_{\S_M} \hat{\m h} \P_{\S_M}
+  = \sum_{s\in\S_M} \f{\ket{s}}{\sqrt{\abs{s}}} \sp{
+    \sum_{\substack{k\in s\\a\in\ZZ_M}}
+    \sum_{\substack{p\in\ZZ_N\\p\notin k}}
+    h_{p k_a} \f{\bra{\sp{k_{a:p}}}}{\sqrt{\abs{\sp{k_{a:p}}}}}
+    + \sum_{\set{a,b}\in\C_M\p{2}} h_{k_ak_b}
+    \f{\bra{\sp{k_{a\lra b}}}}{\sqrt{\abs{\sp{k_{a\lra b}}}}}
+    - h^{\t{D}}_s \f{\bra{s}}{\sqrt{\abs{s}}}},
+\end{align}
+where $h^{\t{D}}_s\equiv h^{\t{D}}_k$ for any $k\in s$, with
+$h^{\t{D}}_k$ defined in \eqref{eq:multi_body_op_diag_element}; and
+$\sp{k}$ denotes the equivalence class of $k$.  The reduced multi-body
 eigenvalue problem is then
 \begin{align}
-  \hat{\m h}_{\S_M} \c \v{\m w} = \Delta \v{\m w}.
+  \P_{\S_M} \hat{\m h} \P_{\S_M} \c \v{\m w} = \Delta \v{\m w}.
 \end{align}
@@
 obey the spatial symmetries as $H_0$, these symmetries generally endow
-$\hat{\m h}$ with a block-diagonal structure. \red{[TODO: finish
-  discussion/derivation]}
+$\hat{\m h}$ with a block-diagonal structure that reduces the
+complexity of the multi-body eigenvalue problem.  We leave this
+reduction to future work.  % todo: perform reduction
 
@@
 $\m X=\p{X_1,X_2}$ with dimensions $\p{M_1,M_2}=\p{2,1}$ and
-$\abs{g}=2$ with $g_{\set{1,2}}=g_{\set{2}}=1$, then
+$\abs{g}=2$ with $g_{\set{1}}=g_{\set{1,2}}=1$, then
 $X_1\p{a,b} X_2\p{b}$ and $X_1\p{b,a} X_2\p{b}$ are distinct
@@
 assigning indices $k=\p{k_1,k_2,\cdots,k_{\abs{g}}}$ to the tensors
-$\m w$ as prescribed by the $\ell$, and summing over all
+$\m w$ as prescribed by $\ell$, and summing over all
 $k\in\D_N\p{\abs{g}}$, while the $\abs{g}$-spin operator
 $\m X\p{\ell}$ is obtained by contracting the operators in $\m X$ in
-accordance with the index assignment $\ell$.  If the tensors $\m w$ or
-operators $\m X$ are all permutationally symmetric, then
-$\m X_{\m w}\p{g}$ factorizes into $\S\p{g} \m w\p{g} \m X\p{g}$,
-where $\m w\p{g}$ and $\m X\p{g}$ are respectively averages of
-$\m w\p{\ell}$ and $\m X\p{\ell}$ over all index assignments
-$\ell\in\L\p{g}$.
+accordance with $\ell$.  If the tensors $\m w$ or operators $\m X$ are
+all permutationally symmetric, then $\m X_{\m w}\p{g}$ factorizes into
+$\S\p{g} \m w\p{g} \m X\p{g}$, where $\m w\p{g}$ and $\m X\p{g}$ are
+respectively averages of $\m w\p{\ell}$ and $\m X\p{\ell}$ over all
+index assignments $\ell\in\L\p{g}$.
 
@@
 where the proportionality is up to an appropriate symmetry factor (in
-this case, $1/2!\times1/2!$; clarified in Section
+this case, $1/2!\times1/2!$, as clarified in Section
 \ref{sec:symmetry_factors}).  Crosses, meanwhile, correspond to
@@
 sequentially eliminating filled dots and crosses until only empty dots
-remain.  The only complication in carrying out this process is that,
-in the case of permutationally symmetric tensors, we have to manually
-keep track of symmetry factors such as the factor of 2 above.
+remain.  The only complication in carrying out this process is that we
+have to manually keep track of symmetry factors, such as the factor of
+2 above.
 
@@
 $F_{\t{old}}$ choices of which dot to move.  More precisely,
-eliminating a cross and moving a dot corresponds to forcing the index
-associated with the cross to be equal to the index associated with the
-filled dot (in the ``old'' region), and there are $F_{\t{old}}$ such
-indices to choose from.
+eliminating a cross and moving a filled dot corresponds to forcing the
+index associated with the cross to be equal to the index associated
+with the filled dot, and there are $F_{\t{old}}$ such indices to
+choose from.
 
@@
 \eqref{eq:sym_prod_diagrams} is systematic enough to be converted into
-an algorithm for execution on a computer.  The last step necessary to
-automate the evaluation of \eqref{eq:sym_prod_diagrams} is the
-expansion of the multi-spin operators $\m X\p{\ell}$ in a basis for
-the PS manifold $\M_0$.  Without loss of generality, we consider a
-single $M$-spin operator $X$; our task is essentially to find
-coefficients of the expansion
+a computer algorithm.  The last step necessary to automate the
+evaluation of \eqref{eq:sym_prod_diagrams} is the expansion of the
+multi-spin operators $\m X\p{\ell}$ in a basis for the PS manifold
+$\M_0$.  Without loss of generality, we consider a single $M$-spin
+operator $X$; our task is essentially to find coefficients of the
+expansion
 \begin{align}

### 2020-03-27 collective_sun/sun_perturbation.tex
added discussion of the algorithm to construct a basis for a multi-body excitation manifold

@@
 $\ket{m}=\ket{m_1,m_2,\cdots,m_n}$ with a definite occupation number
-$m_\mu$ of the single-spin state $\mu$, and $\sum_\mu m_\mu=N$.  In
-the case of $n=2$, for example, the PS manifold $\M_0$ is precisely
-the Dicke manifold\cite{dicke1954coherence} of $N$ qubits, spanned by
-Dicke states $\set{\ket{m_0,m_1}}$ for which $m_0+m_1=N$ and the
-integer $m_0$ ($m_1$) indicates the total occupation number of qubit
-state $\ket{0}$ ($\ket{1}$).  The dimension of the PS manifold $\M_0$
-is determined by the number of ways to assign $N$ spins to $n$ states,
-which is ${N+n-1 \choose n-1}\sim N^{n-1}$.  The energy of any PS
-state $\ket\psi\in\M_0$ with respect to the interaction Hamiltonian
-$H_0$ in \eqref{eq:H_0} is
+$m_\mu$ of the single-spin state $\mu$, and $\sum_\mu m_\mu=N$.  We
+denote the set of all such assignments of $N$ (identical) spins to $n$
+(distinct) states by $\A_n\p{N}$, such that our standard basis for
+$\M_0$ is $\set{\ket{m}:m\in\A_n\p{N}}$.  The dimension of $\M_0$ is
+then the number of such assignments, which is straightforward to
+determine with a standard ``stars and bars'' calculation:
+$\abs{\A_n\p{N}} = {N+n-1 \choose n-1} \sim N^{n-1}$.  In the case of
+$n=2$, for example, the PS manifold $\M_0$ is precisely the Dicke
+manifold\cite{dicke1954coherence} of $N$ qubits, spanned by
+${N+2-1\choose2-1}=N+1$ Dicke states $\ket{m_0,m_1}$ for which
+$m_0+m_1=N$ and the integer $m_0$ ($m_1$) indicates the total
+occupation number of qubit state $\ket{0}$ ($\ket{1}$).  The energy of
+any PS state $\ket\psi\in\M_0$ with respect to the interaction
+Hamiltonian $H_0$ in \eqref{eq:H_0} is
 \begin{align}
@@
   + \m h_{\t{P}}\p{\m w} - \m h_{\t{D}}\p{\m w}.
+  \label{eq:multi_body_op}
 \end{align}
@@
 therefore diagonalizable.  Diagonalizing this operator yields
-eigenvalues $\Delta$ and eigenvectors $\m m_\Delta$ for which
-$\hat{\m h}\p{\m m_\Delta}=\Delta\m m_\Delta$.  Finding eigenvalues
+eigenvalues $\Delta$ and eigenvectors $\m u_\Delta$ for which
+$\hat{\m h}\p{\m u_\Delta}=\Delta\m u_\Delta$.  Finding eigenvalues
 and eigenvectors of $\hat{\m h}$ nominally requires diagonalizing a
@@
 We then denote the a direct sum of all $\M_r$ for $r\le M$ by
-$\ul{\M}_M$, i.e.~$\ul{\M}_M\equiv\bigoplus_{r\le M}\M_r$.  The
-manifold $\ul{\M}_M$ consists of all states accessible by the action
-of $r$-body operators with $r\le M$ on PS states in $\M_0$.
+$\ul\M_M$, i.e.~$\ul\M_M\equiv\bigoplus_{r\le M}\M_r$.  The manifold
+$\ul\M_M$ consists of all states accessible by the action of $r$-body
+operators with $r\le M$ on PS states in $\M_0$.
 
 By construction, the $M$-body operator $X_{\m w}$ generally couples PS
-states in $\M_0$ to asymmetric states in $\ul{\M}_M\setminus\M_0$.  If
+states in $\M_0$ to asymmetric states in $\ul\M_M\setminus\M_0$.  If
 the coupling between $\M_0$ and its complement is weak compared to the
@@
 $\norm{X_{\m w}}\ge\Delta_{\t{gap}}/2$, in which case the operator
-$X_{\m w}$ first couples $\M_0$ to $\ul{\M}_M$, then $\ul{\M}_M$ to
-$\ul{\M}_{2M}$, and so on.  Even if
+$X_{\m w}$ first couples $\M_0$ to $\ul\M_M$, then $\ul\M_M$ to
+$\ul\M_{2M}$, and so on.  Even if
 $\norm{X_{\m w}}>\Delta_{\t{gap}}/2$, however, the energetic penalty
@@
 the dynamics of an initial PS state $\ket\psi\in\M_0$ by projecting
-$X_{\m w}$ onto $\ul{\M}_{r_{\t{max}}}$ for some $r_{\t{max}}\ge M$,
+$X_{\m w}$ onto $\ul\M_{r_{\t{max}}}$ for some $r_{\t{max}}\ge M$,
 resulting in a theory that resembles first-order perturbation theory
-on $\ul{\M}_{r_{\t{max}}}$.  In the perturbative regime
+on $\ul\M_{r_{\t{max}}}$.  In the perturbative regime
 $\norm{X_{\m w}}<\Delta_{\t{gap}}/2$, such a theory captures all
@@
 truncating the Hilbert space of a spin system past
-$\ul{\M}_{r_{\t{max}}}$ is not clear a priori.  Nonetheless,
-simulations restricted to $\ul{\M}_{r_{\t{max}}}$ can be benchmarked
-by examining the maximum populations within all $\M_r$ for
-$r\le r_{\t{max}}$ throughout simulation.  As long as
+$\ul\M_{r_{\t{max}}}$ is not clear a priori.  Nonetheless, simulations
+restricted to $\ul\M_{r_{\t{max}}}$ can be benchmarked by examining
+the maximum populations within all $\M_r$ for $r\le r_{\t{max}}$
+throughout simulation.  As long as
 \begin{enumerate*}
-\item these
-populations fall off with increasing $r$,
+\item these populations fall off with increasing $r$,
 \item and the maximal population within large-$r$ manifolds $\M_r$ is
@@
 
-In order to project an operator $X_{\m w}$ onto
-$\ul{\M}_{r_{\t{max}}}$, we first need to construct a basis for
-$\ul{\M}_{r_{\t{max}}}$, and then compute all matrix elements of
-$X_{\m w}$ in the constructed basis.  Our strategy will be to first
-identify a basis for $\M_0$, and then find, for each
+In order to project an operator $X_{\m w}$ onto $\ul\M_{r_{\t{max}}}$,
+we first need to construct a basis for $\ul\M_{r_{\t{max}}}$, and then
+compute all matrix elements of $X_{\m w}$ in this basis.  Our strategy
+will be to first identify a basis for $\M_0$, and then find, for each
 $r=1,\cdots,r_{\t{max}}$, a suitable set of $r$-body operators
@@
 $\M_r$.  Matrix elements of $X_{\m w}$ with respect to this basis for
-$\ul{\M}_{r_{\t{max}}}$ then take the form
+$\ul\M_{r_{\t{max}}}$ then take the form
 \begin{align}
@@
 
-A suitable basis for $\M_0$ is the set of PS states
-$\ket{m}=\ket{m_1,m_2,\cdots,m_n}$ with a definite occupation number
-$m_\mu$ of the single-spin state $\mu$.
+Denoting all assignments of $N$ (identical) spins to $n$ (distinct)
+states by $\A_n\p{N}$, a suitable basis for $\M_0$ is the set of PS
+states $\set{\ket{m}:m\in\A_n\p{N}}$, where each
+$m=\p{m_1,m_2,\cdots,m_n}$ specifies the occupation number $m_\mu$ of
+the single-spin state $\mu$.  For consistency with the expression of
+matrix elements in \eqref{eq:matrix_element}, for each
+$\ket{m}\in\M_0$ we define $\ket{\1,1,m}\equiv\ket{m}$, with $\1$ the
+identity operator, and we define $\1_1\equiv\1$.  We now need to
+determine the set of $r$-body operators $\set{O}$ and dimension-$r$
+tensors $\set{\m u}$ that define our basis for $\M_r$ with
+$r=1,2,\cdots,r_{\t{max}}$.  As an initial guess for $\set{\m u}$, we
+can choose the set $\set{\m u_\Delta}_{\m h,r}$ of dimension-$r$
+tensors $\m u_\Delta$ that generate eigenstates of $H_0$, i.e.~for
+which $\hat{\m h}\p{\m u_\Delta} = \Delta \m u_\Delta$, with
+$\hat{\m h}$ defined in \eqref{eq:multi_body_op}.  The set of $r$-body
+operators to choose is less obvious, but can be determined by an
+exhaustive search over all $r$-body operators relevant to a particular
+scenario.  Specifically, the purpose of projecting operators onto
+$\col\M_{r_{\t{max}}}$ is to simulate the dynamics induced by some
+operator $\hat\O$ of the form in \eqref{eq:pert}, constructed from a
+set $\O=\set{\p{X,\m w}}$ containing multi-spin operators $X$.  In
+order to find a basis for the {\it relevant} states in $\M_r$, one can
+simply consider the set $\set{O}_{\O,r}$ all $r$-spin operators that
+can be constructed from products of the multi-spin operators $X$ in
+$\O$.
+
+In general, the states $\ket{O,\m u_\Delta,m}$ for all
+$\p{O,\m u_\Delta,m}\in\set{O}_{\O,r}\times\set{\m u_\Delta}_{\m h,r}
+\times\A_n\p{N}$ will form an overcomplete basis for the relevant
+states in $\col\M_r$.  Our last step is therefore to reduce this set
+of states to a minimal basis $\B_r$ for $\M_r$.  To this end, we
+construct each of $\B_1,\B_2,\cdots,\B_{r_{\t{max}}}$ in sequence,
+such that when constructing $\B_r$ we already have a minimal basis
+$\col\B_{r-1}\equiv\bigcup_{s<r-1}\B_s$ for $\col\M_{r-1}$.  To
+construct $\B_r$, we then consider each
+$\p{O,\m u_\Delta,m}\in\set{O}_{\O,r}\times\set{\m u_\Delta}_{\m h,r}
+\times\A_n\p{N}$ one at a time.  If $\ket{O,\m u_\Delta,m}$ is
+orthogonal to all states in $\col\B_{r-1}$, as well as all other
+states we have added to $\B_r$ thus far, then we include it in $\B_r$.
+Otherwise, we discard $\ket{O,\m u_\Delta,m}$.  There are two
+considerations to keep in mind for this construction: first, one
+should always throw out null states $\ket{O,\m u_\Delta,m}=0$, which
+can be identified by a vanishing self-overlap,
+$\bk{O,\m u_\Delta,m|O,\m u_\Delta,m}=0$.  Second, computing overlaps
+to check orthogonality can be computationally intensive.  When
+considering a candidate state $\ket{O,\m u_\Delta,m}$ to add to
+$\B_r$, one can therefore restrict orthogonality checks to other
+states $\ket{Q,\m v_\epsilon,\ell}$ with $\epsilon=\Delta$; if the
+excitation energies $\epsilon\ne\Delta$, then these states are
+guaranteed to be orthogonal.  Any two states with different occupation
+numbers of single-spin states $\mu\in\ZZ_n$ are similarly guaranteed
+to be orthogonal.
 
 \vspace{3cm}
-
-% For consistency with \eqref{eq:matrix_element}, for each state
-% $\ket{m}$ in this basis for $\M_0$ we define
-% $\ket{\1,1,m}\equiv\ket{m}$, with $\1$ the identity operator, and we
-% define $\1_1\equiv\1$.
 
@@
   + \m h_{\t{P}}\p{\m w} - \m h_{\t{D}}\p{\m w},
-  \label{eq:multi_body_op}
 \end{align}
@@
 $\M_0$.  Without loss of generality, we consider a single $M$-spin
-operator $X$; our task is essentially to find coefficients of the
+operator $X$.  Our task is essentially to find coefficients of the
 expansion
@@
   \P_0 X \P_0
-  = \sum_{a,b\in\B_0} \bk{a|X|b} \op{a}{b},
-\end{align}
-where $\set{\ket{a}:a\in\B_0}$ is complete bases for $\M_0$, and the
-choice of $M$ spins on which $X$ acts is arbitrary.  In a system of
-$N$ spins with $n$ states each, a suitable basis for $\M_0$ can be
-labeled by the occupation number $a_\mu$ of each single-spin state,
-namely $a=\p{a_1,a_2,\cdots,a_n}$ with $\sum_\mu a_\mu=N$.  Written
-out explicitly,
+  = \sum_{a,b\in\A_n\p{N}} \bk{a|X|b} \op{a}{b},
+\end{align}
+where $\A_n\p{N}$ is the set of all ways to assign $N$ (identical)
+spins to $n$ (distinct) states, such that for any $a\in\A_n\p{N}$, the
+state $\ket{a}=\ket{a_1,a_2,\cdots,a_n}$ is labeled by the occupation
+number $a_\mu$ of state $\mu$, with $\sum_\mu a_\mu = N$.  Written out
+explicitly,
 \begin{align}
@@
 distinct ways to permute the tensor factors of $\ket{\tilde a}$, and
-enforces $\bk{a|a}=1$.  We denote the set of all valid assignments of
-$N$ (identical) spins to $n$ (distinguishable) states by $\A_n\p{N}$,
-such that the basis $\set{\ket{a}:a\in\A_n\p{N}}$ spans
-permutationally symmetric manifold of $N$ spins.  Identifying this
-basis allows us to expand
+enforces $\bk{a|a}=1$.  Using these states, we can expand
 \begin{align}

### 2020-03-27 collective_sun/sun_perturbation.tex
fixed up some text in the shell model section

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Breaking SU($n$) symmetry (perturbatively)}
-\label{eq:pert_theory}
+\section{Breaking SU($n$) symmetry}
 
@@
 matrix of dimensions $\abs{\D_N\p{M}}\times\abs{\D_N\p{M}}$, but
-spatial symmetries of $H_0$ and $X_{\m w}$ such as translational
-invariance or isotropy can drastically reduce the complexity of this
+spatial symmetries of $H_0$ and $X_{\m w}$, such as translational
+invariance or isotropy, can drastically reduce the complexity of this
 eigenvalue problem.  We discuss these reductions in Appendix
@@
 \section{Beyond ground-state perturbation theory}
-\label{eq:excited_states}
 
@@
 states by $\A_n\p{N}$, a suitable basis for $\M_0$ is the set of PS
-states $\set{\ket{m}:m\in\A_n\p{N}}$, where each
-$m=\p{m_1,m_2,\cdots,m_n}$ specifies the occupation number $m_\mu$ of
-the single-spin state $\mu$.  For consistency with the expression of
-matrix elements in \eqref{eq:matrix_element}, for each
-$\ket{m}\in\M_0$ we define $\ket{\1,1,m}\equiv\ket{m}$, with $\1$ the
-identity operator, and we define $\1_1\equiv\1$.  We now need to
-determine the set of $r$-body operators $\set{O}$ and dimension-$r$
-tensors $\set{\m u}$ that define our basis for $\M_r$ with
-$r=1,2,\cdots,r_{\t{max}}$.  As an initial guess for $\set{\m u}$, we
+states $\set{\ket{m}:m\in\A_n\p{N}}$, where $m=\p{m_1,m_2,\cdots,m_n}$
+specifies the occupation number $m_\mu$ of each single-spin state
+$\mu$.  For consistency with the expression of matrix elements in
+\eqref{eq:matrix_element}, for each $\ket{m}\in\M_0$ we define
+$\ket{\1,1,m}\equiv\ket{m}$, with $\1$ the identity operator, and we
+define $\1_1\equiv\1$.  We now need to determine the set of $r$-body
+operators $\set{O}$ and dimension-$r$ tensors $\set{\m u}$ that define
+our basis for each $\M_r$.  As an initial guess for $\set{\m u}$, we
 can choose the set $\set{\m u_\Delta}_{\m h,r}$ of dimension-$r$
@@
 such that when constructing $\B_r$ we already have a minimal basis
-$\col\B_{r-1}\equiv\bigcup_{s<r-1}\B_s$ for $\col\M_{r-1}$.  To
+$\col\B_{r-1}\equiv\bigcup_{s<r}\B_s$ for $\col\M_{r-1}$.  To
 construct $\B_r$, we then consider each
@@
 $\bk{O,\m u_\Delta,m|O,\m u_\Delta,m}=0$.  Second, computing overlaps
-to check orthogonality can be computationally intensive.  When
+to check for orthogonality can be computationally intensive.  When
 considering a candidate state $\ket{O,\m u_\Delta,m}$ to add to
@@
 
-\vspace{3cm}
-
-\red{[TODO: finish section]}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{The XXZ model}
+
+\red{[TODO: write section]}
 

### 2020-03-27 multi_body_interactions/overlap_integrals.tex
adding old notes on computing lattice overlap integrals

\documentclass[aps,notitlepage,nofootinbib,11pt]{revtex4-1}

% linking references
\usepackage{hyperref}
\hypersetup{
  breaklinks=true,
  colorlinks=true,
  linkcolor=blue,
  filecolor=magenta,
  urlcolor=cyan,
}

%%% paragraph indentation and spacing
\setlength{\parindent}{0cm} % don't indent new paragraphs...
\parskip 6pt % ... place a space between paragraphs instead
\frenchspacing % only one space after periods

%%% header / footer
\usepackage{fancyhdr} % easier header and footer management
\pagestyle{fancy} % page formatting style
\fancyhf{}
\renewcommand{\headrulewidth}{0pt} % remove horizontal line in header
\usepackage{lastpage} % for referencing last page
\cfoot{\thepage~of \pageref{LastPage}} % "x of y" page labeling

% figures
\usepackage{hyperref} % for linking references
\usepackage{graphicx,float} % for figures
\usepackage{grffile} % help latex properly identify figure extensions
\graphicspath{{./figures/}} % set path for all graphics
\usepackage[caption=false]{subfig} % subfigures
\newcommand{\sref}[1]{\protect\subref{#1}}

%%% symbols, notations, etc.
\usepackage{physics,bm,braket,amssymb} % physics and math packages
\usepackage{accents} % for resolving some accent (e.g. tilde) issues
\renewcommand{\t}{\text} % text in math mode
\newcommand{\f}[2]{\dfrac{#1}{#2}} % shorthand
\newcommand{\p}[1]{\left(#1\right)} % parenthesis
\renewcommand{\sp}[1]{\left[#1\right]} % square parenthesis
\renewcommand{\set}[1]{\left\{#1\right\}} % curly parenthesis
\renewcommand{\v}{\bm} % bold vectors
\newcommand{\uv}[1]{\hat{\v{#1}}} % unit vectors
\renewcommand{\d}{\partial} % partial d
\renewcommand{\c}{\cdot} % inner product
\newcommand{\bk}{\Braket} % shorthand for braket notation

\usepackage{dsfont}
\newcommand{\1}{\mathds{1}}

\newcommand{\E}{\mathcal E}

\usepackage{listings}

\begin{document}

\title{Efficient calculation of Wannier functions and their overlap
  integrals}

\author{Michael A. Perlin}

\maketitle
\thispagestyle{fancy}

Solutions to the Mathieu equation (i.e. spatial eigenfunctions of an
optical lattice, or Bloch functions) with quasi-momentum $q$ in band
$n$ take the form
\begin{align}
  \tilde\phi_{qn}\p{x}
  = e^{iqx} \sum_{k=-\infty}^\infty c_{qk}^{(n)}e^{2ikx}
\end{align}
where all coefficients $c_{qk}^{(n)}\in\mathbb R$. Without loss of
generality, we will only consider on-site (Wannier) functions
$\phi_n\p{x}$ for band $n$ on a lattice site centered at $x=0$, which
are given by the discrete Fourier transform of the Bloch functions:
\begin{align}
  \phi_n\p{x}
  = \sum_q \tilde\phi_{qn}\p{x}
  = \sum_{q,k} c_{qk}^{(n)} e^{i\p{2k+q}x}
  = \sum_{q,k} c_{qk}^{(n)} \E_{qk}\p{x},
  \label{eq:wannier_first}
\end{align}
where $\E_{qk}\p{x} \equiv e^{i\p{2k+q}x}$. In practice, when
computing \eqref{eq:wannier_first} we must choose a finite number $Q$
of allowed quasi-momenta $q$ and a finite number $K$ of allowed values
of $k$ (centered on 0). We can then construct the $Q\times K$ matrices
$\v c_n\equiv \sum_{q,k}c_{qk}^{(n)}\uv q\otimes\uv k$ and
$\v\E\p{x}\equiv \sum_{q,k}\E_{qk}\p{x}\uv q\otimes\uv k$, and rewrite
\eqref{eq:wannier_first} as
\begin{align}
  \phi_n\p{x} = \v c_n \c \v\E\p{x}.
  \label{eq:wannier}
\end{align}
Constructing these matrices and computing their scalar product is
substantially faster to perform numerically than the double sum in
\eqref{eq:wannier_first}.

The Wannier functions in \eqref{eq:wannier} appear in e.g. the
(one-dimensional) four-function overlap integral
\begin{align}
  F_n
  \equiv \int dx~ \phi_n\p{x}^*\phi_0\p{x}^*\phi_0\p{x}\phi_0\p{x}.
  \label{eq:F}
\end{align}
As $\phi_n\p{x}$ has the same parity (i.e. even/odd) as $n$, the
entire integrand in \eqref{eq:F} likewise has the same parity as
$n$. It follows that $F_n$ vanishes if $n$ is odd. If $n$ is even,
then the integrand in \eqref{eq:F} is even, which means that we need
only integrate over half of the lattice. Furthermore, knowing that
$\phi_0\p{x}$ and $F_n$ are real\footnote{We will not spend time
  justifying this claim here, but can do so upon request.}, we can
generally make the replacement
$\v\E\p{x}\to\Re\sp{\v\E\p{x}}\equiv\v\E^R\p{x}$ for the purposes of
computing \eqref{eq:F}. For a lattice of length $L$, the integral in
\eqref{eq:F} then becomes
\begin{align}
  F_n
  = 2\int_0^{L/2} dx~ \Re\sp{\phi_n\p{x}} \phi_0\p{x}^3
  = 2\int_0^{L/2} dx~
  \sp{\v c_n\c\v\E^R\p{x}} \sp{\v c_0\c\v\E^R\p{x}}^3.
  \label{eq:F2}
\end{align}

In order to reduce the number of operations necessary to compute the
overlaps $F_n$, we should first numerically solve the Mathieu equation
to get the coefficients $c_{qn}^{(k)}$ for every quasi-momentum $q$ up
to some cutoff in band ($n$) and Fourier order ($k$). These
coefficients should be stored in a three-dimensional array which can
be sliced to extract the matrix $\v c_n$ from memory. Recycling the
coefficients $c_{qn}^{(k)}$ in this manner provides an enormous
speedup in the computation of $F_n$, as we would otherwise need to
re-compute all of these coefficients for every instance of the
integrand (as occurs when e.g. using Mathematica's Mathieu function
methods to compute $F_n$). Computing $\v\E^R\p{x}$ only once for each
instance of the integrand and recycling it for $\phi_n\p{x}$ and
$\phi_0\p{x}$ also also saves us an additional factor of $\sim2$ in
computation time.

Lastly, there is another trick we can use to efficiently compute the
matrices $\v\E\p{x}$. Rather than looping over all $q$ and $k$
individually to compute all elements of $\v\E\p{x}$, it is
substantially faster to compute the array $\sum_q e^{iqx}\uv q$ and
replicate it $K$ times; and likewise compute $\sum_k e^{2ikx}\uv k$
and replicate it $Q$ times. We can then multiply these matrices
element-wise to get $\v\E^R\p{x}$. Implemented in Python using the
NumPy library, this procedure would look something like:
\begin{lstlisting}[language=python,basicstyle=\small]
  q_phases = numpy.matlib.repmat(numpy.exp(i * quasi_momenta * x), K, 1).T
  k_phases = numpy.matlib.repmat(numpy.exp(i * k_values * x), Q, 1)
  E_x = q_phases * k_phases
\end{lstlisting}
where ``quasi\_momenta'' and ``k\_values'' are respectively the NumPy
arrays $\sum_q q\uv q$ and $\sum_k k\uv k$.

\end{document}

### 2020-04-01 collective_sun/sun_perturbation.tex
fixed a minor error

@@
 \newcommand{\floor}[1]{\lfloor{#1}\rfloor}
-
-\def\obra#1{\mathinner{({#1}|}}
-\def\oket#1{\mathinner{|{#1})}}
-\def\obk#1{\mathinner{({#1})}}
-\def\oop#1#2{\oket{#1}\!\obra{#2}}
 
@@
 dynamics induced by non-perturbative additions $\hat\O$ to the SU($n$)
-Heisenberg model, with operator norms
-$\norm*{\hat\O}\sim\Delta_{\t{gap}}$.
+Heisenberg model, with operator norms $\norm*{\hat\O}$ comparable to
+the spectral gap $\Delta_{\t{gap}}$ of $H_0$.
 
@@
   \bar w \equiv \f1{\abs{\D_N\p{M}}}
-  \sum_{k\in\D_N\p{M}} w_{\sigma\p{k}},
+  \sum_{k\in\D_N\p{M}} w_k,
   &&

### 2020-04-01 collective_sun/sun_perturbation.tex
add sketch of multi-body excitation manifolds

@@
 \renewcommand*{\thefootnote}{\alph{footnote}}
+
+% for energy diagrams, etc.
+\usepackage{tikz}
+\usetikzlibrary{arrows,math}
+\tikzstyle{every picture}+=[remember picture]
 
@@
 $\ul\M_M$ consists of all states accessible by the action of $r$-body
-operators with $r\le M$ on PS states in $\M_0$.
+operators with $r\le M$ on PS states in $\M_0$.  These definitions are
+sketched out in Figure \ref{fig:manifold_sketch}.
+
+\begin{figure}
+  \centering
+  \begin{tikzpicture}
+    \draw[ultra thick] (0,0) -- node[above] {$\M_0$} ++(10em,0);
+    \draw[ultra thick] (0,3em) -- node[above] {$\M_1$} ++(10em,0);
+    \draw[ultra thick] (0,2*3em) -- node[above] {$\M_2$} ++(10em,0);
+    \draw[->] (-.2em,0) to [bend left] (-.2em,3em);
+    \node (X) at (.2em,0.5*3em) {$\hat\O_1$};
+    \draw[->] (-.4em,0) to [bend left] (-.2em,2*3em);
+    \node (XY) at (-2em,3em) {$\hat\O_2$};
+    \draw (10em+.2em,-.2em) -- (10em+.4em,-.2em)
+    -- (10em+.4em,3em+.2em) -- (10em+.2em,3em+.2em);
+    \node (cM1) at (10em+1.5em,0.5*3em) {$\col\M_1$};
+    \draw (10em+2.4em+.2em,-.4em) -- (10em+2.4em+.4em,-.4em)
+    -- (10em+2.4em+.4em,6em+.4em) -- (10em+2.4em+.2em,6em+.4em);
+    \node (cM1) at (10em+2.4em+1.5em,3em) {$\col\M_2$};
+  \end{tikzpicture}
+  \caption{Schematic of low-lying multi-body manifolds.  Starting with
+    the permutationally-state manifold $\M_0$, the single-body
+    excitation manifold $\M_1$ consists of states accessible via
+    single-body perturbations $\hat\O_1$.  The two-body excitation
+    manifold $\hat\O_2$ then consists of states accessible via
+    two-body perturbations, but not single-body perturbations.  The
+    joint manifold $\col\M_M$ consists of all states accessible by
+    operators $r$-body operators $\O_r$ with $r\le M$.}
+  \label{fig:manifold_sketch}
+\end{figure}
 
 By construction, the $M$-body operator $X_{\m w}$ generally couples PS
-states in $\M_0$ to asymmetric states in $\ul\M_M\setminus\M_0$.  If
-the coupling between $\M_0$ and its complement is weak compared to the
-spectral gap $\Delta_{\t{gap}}$ of $H_0$, then we can treat the effect
-of $X_{\m w}$ perturbatively and derive an effective Hamiltonian
+states in $\M_0$ to asymmetric states in $\ul\M_M$.  If the coupling
+between $\M_0$ and its complement is weak compared to the spectral gap
+$\Delta_{\t{gap}}$ of $H_0$, then we can treat the effect of
+$X_{\m w}$ perturbatively and derive an effective Hamiltonian
 $H_{\t{eff}}$ on $\M_0$.  This ground-state perturbation theory

### 2020-04-02 collective_sun/sun_perturbation.tex
added some todo statements drastically simplified the section on perturbation theory started an introduction

@@
 
+% todo: solve collective model; Casmir operators?
+
 \title{SU($n$) spin models near the Heisenberg critical point}%
@@
 
-\red{TODO: write an introduction}
+SU($n$) symmetries play an important role in contemporary physics.
+Underpinning much of high energy physics, the celebrated SU($n$) gauge
+theory known as Yang-Mills theory is central to our understanding of
+the electroweak and strong forces.  Extensions of Yang-Mills and
+SU($n$) symmetry feature in the most well-studied examples of
+holographic duality\cite{maldacena1999largen} and the connection
+between entanglement and gravity\cite{ryu2006holographic} through the
+anti-de Sitter / conformal field theory (AdS/CFT) correspondence.  In
+a condensed matter setting, SU(2) appears ubiquitously as a symmetry
+of the Hubbard model, with important consequences for the study of
+quantum magnetism and high temperature
+superconductivity\cite{lee2006doping}.  The extension of SU(2) spin
+models to the case of SU($n$) has led to predictions of exotic new
+phases of matter\cite{hermele2009mott, hermele2011topological,
+  chen2016syntheticgaugefield, nataf2016chiral}, as well as the
+potential to perform universal topological quantum
+computation\cite{freedman2004class, nayak2008nonabelian}.
+Furthermore, the introduction of disorder into SU($n$) spin models has
+opened avenues for studying quantum chaos and information
+scrambling\cite{sachdev1993gapless, bentsen2019integrable}.
+
+Last but not least, SU($n$) symmetries appear naturally in the physics
+of atomic, molecular, and optical (AMO)
+systems\cite{beverland2016realizing, cazalilla2014ultracold,
+  gorshkov2010twoorbital, scazza2014observation, taie2012su,
+  zhang2014spectroscopic, goban2018emergence, perlin2019effective},
+enabling their experimental study with an exquisite level of control.
+These symmetries typically appear through the independence of
+inter-atomic interaction parameters on the $n$ nuclear spin states of
+the individual atoms involved, with e.g.~$n=10$ in the case of
+${}^{87}$Sr.
+
+\red{[finish introduction]}
+
+\vspace{3cm}
 
@@
 \section{The SU($n$) Heisenberg model}
+
+% todo: give example of SU(2)
 
@@
 
+% todo: simplify, give examples at every step
+% todo: consider special case before a general one
+% todo: new symbol for X_w, e.g. O(X,w)
+
 In order to induce non-trivial dynamics for initially PS states
@@
 is included in \eqref{eq:pert} by convention to average over the $M!$
-terms addressing any given set of spins.  Due to the permutational
-symmetry enforced by the interaction Hamiltonian $H_0$, we will find
-it convenient to decompose $\D_N\p{M}=\pi\p{M}\times\C_N\p{M}$, where
-$\pi\p{M}$ is the permutation group of $\ZZ_M$, and
-\begin{align}
-  \C_N\p{M} \equiv \set{ k \subset \ZZ_N: \abs{k} = M },
-  \label{eq:choices}
-\end{align}
-is the set of all subsets (``choices'') of $M$ elements from $\ZZ_N$.
-We can then expand
-\begin{align}
-  X_{\m w} = \f1{M!} \sum_{\sigma\in\pi\p{M}} \sum_{k\in\C_N\p{M}}
-  w_{\sigma\p{k}} X_{\sigma\p{k}}.
-\end{align}
-To simplify language and notation, we will assume throughout this work
-that $\hat\O$ is homogeneous in the sense that $M$ is the same for all
+terms addressing any given set of spins.  To simplify language and
+notation, we will assume throughout this work that $\hat\O$ is
+homogeneous in the sense that $M$ is the same for all
 $\p{X,\m w}\in\O$.  A generalization to the case of inhomogeneous
@@
 correspond to two-body interactions that deviate from $H_0$, such as
-$s_\z\otimes s_\z$ interactions in an XXZ model\red{[TODO:cite]} near
-the Heisenberg critical point.  While we would be hard-pressed to find
-a physical implementation of multi-body spin interactions with $M>2$,
-considering this case will be necessary to construct low-lying
-eigenstates of $H_0$, which will in turn enable simulating the
-dynamics induced by non-perturbative additions $\hat\O$ to the SU($n$)
-Heisenberg model, with operator norms $\norm*{\hat\O}$ comparable to
-the spectral gap $\Delta_{\t{gap}}$ of $H_0$.
+$s_\z\otimes s_\z$ interactions in an XXZ model near the Heisenberg
+critical point.  While we would be hard-pressed to find a physical
+implementation of multi-body spin interactions with $M>2$, considering
+this case will be necessary to construct low-lying eigenstates of
+$H_0$, which will in turn enable simulating the dynamics induced by
+non-perturbative additions $\hat\O$ to the SU($n$) Heisenberg model,
+with operator norms $\norm*{\hat\O}$ comparable to the spectral gap
+$\Delta_{\t{gap}}$ of $H_0$.
 
@@
 projector onto $\M_0$ itself, and $\E_0=\M_0$. The first order
-effective Hamiltonian $H_{\t{eff}}^{(1)}$ is straightforward to
-simplify using the permutational symmetry of $\M_0$:
+effective Hamiltonian $H_{\t{eff}}^{(1)}$ simply projects the
+perturbation $\hat\O$ onto the permutationally-symmetric manifold
+$\M_0$, and takes the form
 \begin{align}
@@
 coefficient $\bar w$ is the average of all $w_k$; and the collective
-operator $\col{X}$ is a permutation-averaged sum of all $X_k$, i.e.
+operator $\col{X}$ is a sum of all $X_k$ averaged over permutations of
+any particular choice of spins $k\in\D_N\p{M}$.  To write out the
+collective operator $\col{X}$ explicitly, we define the set \begin{align}
+  \C_N\p{M} \equiv \set{ k \subset \ZZ_N: \abs{k} = M },
+  \label{eq:choices}
+\end{align}
+is the set of all subsets (``choices'') of $M$ elements from $\ZZ_N$,
+and the set $\pi\p{M}$ of all permutations of $M$ elements, in terms
+of which
 \begin{align}
@@
 \end{align}
-with $\abs{\D_N\p{M}}=\prod_{j=0}^{M-1}\p{N-j}$ and
-$\abs{\pi\p{M}}=M!$.  An immediate consequence of \eqref{eq:H_eff_1}
-is the fact that perturbing SU(2)-symmetric interactions by two-body
-$s_\z\otimes s_\z$ interactions yields the one-axis twisting (OAT)
+where $\abs{\D_N\p{M}}=\prod_{j=0}^{M-1}\p{N-j}$ and
+$\abs{\pi\p{M}}=M!$.  Here $k\in\C_N\p{M}$ defines which spins
+$X_{\sigma\p{k}}$ acts on, while $\sigma\in\pi\p{M}$ determines which
+tensor factor of $X$ each of these spins is associated with.  An
+immediate consequence of \eqref{eq:H_eff_1} is the fact that
+perturbing SU(2)-symmetric interactions by two-body $s_\z\otimes s_\z$
+interactions yields the one-axis twisting (OAT)
 Hamiltonian\cite{kitagawa1993squeezed, ma2011quantum}
@@
 off-diagonal with respect to the ground- and excited-state manifolds
-$\E_0$ and $\E_\Delta$.  To decompose $X_{\m w}$ into components that
-generate states of definite excitation energy when acting on PS states
-$\M_0$, we pick an arbitrary state $\ket\psi\in\M_0$ and expand (see
-Appendix \ref{sec:eigenstates})
-\begin{align}
-  H_0 X_{\m w} \ket\psi
-  = E_0 X_{\m w} \ket\psi + X_{\hat{\m h}\p{\m w}} \ket\psi,
-  \label{eq:diagnosis}
-\end{align}
-where $\hat{\m h}$ is a linear map on the space of dimension-$M$
-tensors, determined entirely by the couplings $h_{pq}$ in $H_0$.  The
-image $\hat{\m h}\p{\m w}$ of $\m w$ under $\hat{\m h}$ is then itself
-a dimension-$M$ tensor that defines the $M$-body operator
-$X_{\hat{\m h}\p{\m w}}$.  To write out $\hat{\m h}\p{\m w}$
-explicitly, we first define a matrix
-$\m h\equiv\sum_{p,q\in\ZZ_N} h_{pq} \op{p}{q}$ of all couplings in
-$H_0$ by enforcing that $h_{pq}=h_{qp}$ and $h_{pp}=0$.  We denote a
-contraction of $\m h$ with the $a$-th index of $\m w$ by
-$\m h_a\p{\m w}$, such that the $k$-th entry of $\m h_a\p{\m w}$ is
-\begin{align}
-  \m h_a\p{\m w}_k
-  \equiv \sum_{\substack{p\in\ZZ_N\\p\notin k}}
-  h_{pk_a} w_{k_1 k_2 \cdots k_{a-1} p k_{a+1} \cdots k_M}
-  = \sum_{\substack{p\in\ZZ_N\\p\notin k}} h_{pk_a} w_{k_{a:p}},
-\end{align}
-where if $k=\p{k_1,k_2,\cdots,k_M}$, then $k_{a:p}$ is the same list
-but with the $a$-th entry $k_a$ replaced by $p$.  We similarly define
-a tensor $\m h_{\t{P}}\p{\m w}$ with entries
-\begin{align}
-  \m h_{\t{P}}\p{\m w}_k
-  \equiv \sum_{\set{a,b}\in\C_M\p{2}} h_{k_ak_b} w_{k_{a\lra b}},
-\end{align}
-where $k_{a\lra b}$ is the same as $k$ but with the placement of $k_a$
-and $k_b$ swapped.  Finally, we define a tensor $\m h_{\t{D}}\p{\m w}$
-with entries
-\begin{align}
-  \m h_{\t{D}}\p{\m w}_k \equiv h^{\t{D}}_k w_k,
-  &&
-  h^{\t{D}}_k \equiv \sum_{p\in k} h_p
-  - \sum_{\set{a,b}\in\C_M\p{2}} h_{k_ak_b},
-  &&
-  h_p \equiv \sum_{q\in\ZZ_N} h_{pq}.
-\end{align}
-These definitions allow us to expand
-\begin{align}
-  \hat{\m h}\p{\m w}
-  \equiv \sum_{a\in\ZZ_M} \m h_a\p{\m w}
-  + \m h_{\t{P}}\p{\m w} - \m h_{\t{D}}\p{\m w}.
-  \label{eq:multi_body_op}
-\end{align}
-Viewed an operator on the vector space of dimension-$M$ tensors, the
-map $\hat{\m h}$ is real and symmetric under transposition, and is
-therefore diagonalizable.  Diagonalizing this operator yields
-eigenvalues $\Delta$ and eigenvectors $\m u_\Delta$ for which
-$\hat{\m h}\p{\m u_\Delta}=\Delta\m u_\Delta$.  Finding eigenvalues
-and eigenvectors of $\hat{\m h}$ nominally requires diagonalizing a
-matrix of dimensions $\abs{\D_N\p{M}}\times\abs{\D_N\p{M}}$, but
-spatial symmetries of $H_0$ and $X_{\m w}$, such as translational
-invariance or isotropy, can drastically reduce the complexity of this
-eigenvalue problem.  We discuss these reductions in Appendix
-\ref{sec:symmetries}.
-
-Defining $\m w_\Delta$ to be the projection of the tensor $\m w$ onto
-the $\Delta$-eigenspace of $\hat{\m h}$, we can expand
-\begin{align}
-  \m w = \sum_\Delta \m w_\Delta,
-  &&
-  \hat{\m h}\p{\m w} = \sum_\Delta \Delta \m w_\Delta,
-\end{align}
-and in turn
-\begin{align}
-  X_{\m w} = \sum_\Delta X_{\m w_\Delta},
-  &&
-  X_{\hat{\m h}\p{\m w}} = \sum_\Delta \Delta X_{\m w_\Delta}.
-\end{align}
-Substituting these expansions into \eqref{eq:diagnosis}, we find
-\begin{align}
-  H_0 X_{\m w} \ket\psi
-  = \sum_\Delta\p{E_0 + \Delta} X_{\m w_\Delta} \ket\psi,
-\end{align}
-which implies that the action of $X_{\m w_\Delta}$ on a PS state
-$\ket\psi\in\M_0$ generates a state with definite interaction energy
-$E_0+\Delta$.  It follows that
-$\P_\Delta X_{\m w} \P_0 = \P_\Delta X_{\m w_\Delta} \P_0 =
-X_{\m{w}_\Delta} \P_0$, so
+$\E_0$ and $\E_\Delta$.  We therefore need to decompose $X_{\m w}$
+into components that generate states of definite excitation energy
+when acting on PS states $\ket\psi\in\M_0$.  The SU($n$) symmetry of
+$H_0$ enables such a decomposition of the form
+\begin{align}
+  H_0 X_{\m w}\ket\psi
+  = \sum_\Delta \p{E_0+\Delta} X_{\m w_\Delta} \ket\psi,
+\end{align}
+where, thinking of the tensor $\m w$ as a $\abs{\D_N\p{M}}$-component
+vector, the tensor $\m w_\Delta$ can be found found by diagonalizing a
+matrix $\hat{\m h}$ of dimensions
+$\abs{\D_N\p{M}}\times\abs{\D_N\p{M}}\sim N^M\times N^M$, and
+projecting $\m w$ onto the eigenspace of $\hat{\m h}$ with eigenvalue
+$\Delta$ (see Appendix \ref{sec:eigenstates}).  The matrix
+$\hat{\m h}$ is determined entirely by the couplings couplings
+$h_{pq}$ of $H_0$.  While the computational cost of diagonalizing
+$\hat{\m h}$ is nominally $O\p{N^{3M}}$, spatial symmetries of $H_0$
+and $X_{\m w}$ such as translational invariance or isotropy can
+drastically reduce the complexity of this eigenvalue problem.  We
+discuss this reduction in Appendix \ref{sec:symmetries}.
+
+Equipped with the decomposition of $X_{\m w}$ into terms that generate
+states of definite excitation energy, we can expand
 \begin{align}
@@
 \section{Beyond ground-state perturbation theory}
+
+% give examples of these manifolds in SU(2)
 
@@
   \end{tikzpicture}
-  \caption{Schematic of low-lying multi-body manifolds.  Starting with
-    the permutationally-state manifold $\M_0$, the single-body
-    excitation manifold $\M_1$ consists of states accessible via
-    single-body perturbations $\hat\O_1$.  The two-body excitation
-    manifold $\hat\O_2$ then consists of states accessible via
-    two-body perturbations, but not single-body perturbations.  The
-    joint manifold $\col\M_M$ consists of all states accessible by
+  \caption{Schematic of low-lying multi-body excitation manifolds.
+    Starting with the permutationally-state manifold $\M_0$, the
+    single-body excitation manifold $\M_1$ consists of states
+    accessible via single-body perturbations $\hat\O_1$.  The two-body
+    excitation manifold $\hat\O_2$ then consists of states accessible
+    via two-body perturbations, but not single-body perturbations.
+    The joint manifold $\col\M_M$ consists of all states accessible by
     operators $r$-body operators $\O_r$ with $r\le M$.}
@@
 can choose the set $\set{\m u_\Delta}_{\m h,r}$ of dimension-$r$
-tensors $\m u_\Delta$ that generate eigenstates of $H_0$, i.e.~for
-which $\hat{\m h}\p{\m u_\Delta} = \Delta \m u_\Delta$, with
-$\hat{\m h}$ defined in \eqref{eq:multi_body_op}.  The set of $r$-body
-operators to choose is less obvious, but can be determined by an
-exhaustive search over all $r$-body operators relevant to a particular
-scenario.  Specifically, the purpose of projecting operators onto
-$\col\M_{r_{\t{max}}}$ is to simulate the dynamics induced by some
-operator $\hat\O$ of the form in \eqref{eq:pert}, constructed from a
-set $\O=\set{\p{X,\m w}}$ containing multi-spin operators $X$.  In
-order to find a basis for the {\it relevant} states in $\M_r$, one can
-simply consider the set $\set{O}_{\O,r}$ all $r$-spin operators that
-can be constructed from products of the multi-spin operators $X$ in
-$\O$.
+tensors $\m u_\Delta$ that generate eigenstates of $H_0$.  The set of
+$r$-body operators to choose is less obvious, but can be determined by
+an exhaustive search over all $r$-body operators relevant to a
+particular scenario.  Specifically, the purpose of projecting
+operators onto $\col\M_{r_{\t{max}}}$ is to simulate the dynamics
+induced by some operator $\hat\O$ of the form in \eqref{eq:pert},
+constructed from a set $\O=\set{\p{X,\m w}}$ containing multi-spin
+operators $X$.  In order to find a basis for the {\it relevant} states
+in $\M_r$, one can simply consider the set $\set{O}_{\O,r}$ all
+$r$-spin operators that can be constructed from products of the
+multi-spin operators $X$ in $\O$.
 
@@
 \label{sec:spectral_gap}
+% todo: make a sub subsection of B.1
 
@@
 that e.g.~$\ket{k_{a\lra b}}=\ket{k}$.  These replacements reduce the
-size of $\hat{\m h}$ from $N!_M\times N!_M$ to
-${N\choose M}\times{N\choose M}$, where
-$N!_M=M!\times{N\choose M}=\prod_{j=0}^{M-1}\p{N-j}$ denotes a falling
-factorial.  We discuss further reduction of the eigenvalue problem in
-\eqref{eq:multi_body_problem} by use of additional symmetries in
-Section \ref{sec:symmetries}.
-
-Some special cases: if $M=1$, then $h^{\t{D}}_k=h_k$, so
+size of $\hat{\m h}$ from $\abs{\D_N\p{M}}\times\abs{\D_N\p{M}}$ to
+$\abs{\C_N\p{M}}\times\abs{\C_N\p{M}}$, where
+$\abs{\D_N\p{M}}=M!\times{N\choose M}=\prod_{j=0}^{M-1}\p{N-j}$ and
+$\abs{\C_N\p{M}}={N\choose M}$.  We discuss further reduction of the
+eigenvalue problem in \eqref{eq:multi_body_problem} by use of
+additional symmetries in Section \ref{sec:symmetries}.
+
+Some special cases of practical interest: if $M=1$, then
+$h^{\t{D}}_k=h_k$, so
 \begin{align}

### 2020-04-03 collective_sun/sun_perturbation.tex
add text to introduction

@@
 quantum magnetism and high temperature
-superconductivity\cite{lee2006doping}.  The extension of SU(2) spin
-models to the case of SU($n$) has led to predictions of exotic new
-phases of matter\cite{hermele2009mott, hermele2011topological,
-  chen2016syntheticgaugefield, nataf2016chiral}, as well as the
-potential to perform universal topological quantum
-computation\cite{freedman2004class, nayak2008nonabelian}.
-Furthermore, the introduction of disorder into SU($n$) spin models has
-opened avenues for studying quantum chaos and information
+superconductivity\cite{lee2006doping}.  The extension of the SU(2)
+Hubbard model to SU($n$) has led to predictions of exotic phases of
+matter such as chiral spin liquids\cite{hermele2009mott,
+  hermele2011topological, chen2016syntheticgaugefield,
+  nataf2016chiral}, as well as the potential to perform universal
+topological quantum computation\cite{freedman2004class,
+  nayak2008nonabelian}.  Furthermore, the consideration of disordered
+SU($n$) spin models has opened analytically tractable avenues for
+studying quantum chaos and information
 scrambling\cite{sachdev1993gapless, bentsen2019integrable}.
@@
 of atomic, molecular, and optical (AMO)
-systems\cite{beverland2016realizing, cazalilla2014ultracold,
-  gorshkov2010twoorbital, scazza2014observation, taie2012su,
-  zhang2014spectroscopic, goban2018emergence, perlin2019effective},
-enabling their experimental study with an exquisite level of control.
-These symmetries typically appear through the independence of
-inter-atomic interaction parameters on the $n$ nuclear spin states of
-the individual atoms involved, with e.g.~$n=10$ in the case of
-${}^{87}$Sr.
+systems\cite{gorshkov2010twoorbital, beverland2016realizing,
+  cazalilla2014ultracold, taie2012su, hofrichter2016direct,
+  cappellini2014direct, scazza2014observation, zhang2014spectroscopic,
+  goban2018emergence, perlin2019effective} through the independence of
+atomic orbital and interaction parameters on the $n$ nuclear spin
+states of individual atoms, with e.g.~$n=10$ in the case of
+${}^{87}$Sr.  As a result, experiments can directly probe the SU($n$)
+Hubbard model, leading to experimental observations of SU($n$) Hubbard
+phases and phase transitions\cite{taie2012su, hofrichter2016direct},
+two-orbital SU($n$) magnetism\cite{cappellini2014direct,
+  scazza2014observation, zhang2014spectroscopic}, and multi-body
+SU($n$)-symmetric interactions\cite{goban2018emergence,
+  perlin2019effective}.  The tremendous theoretical significance of
+SU($n$) symmetries makes it all the more exciting that they appear in
+experimental platforms with exquisite degrees of microscopic control,
+thereby enabling the detailed study of these symmetries' consequences
+for fundamental questions in physics, as well as their use in
+practical technological applications.  SU(2)-symmetric interactions,
+for example, can play an important role in the development of quantum
+sensors that make use of entanglement to surpass classical limits on
+measurement precision\cite{he2019engineering}\red{[cite XXZ squeezing
+  paper]}.  The exciting prospect of similarly exploiting larger
+SU($n$) symmetries to achieve a technological advantage is still an
+unexplored avenue with enormous potential.
 
 \red{[finish introduction]}
+
+\vfill
 

### 2020-04-03 collective_sun/sun_perturbation.tex
finish first draft of introduction

@@
 
-\title{SU($n$) spin models near the Heisenberg critical point}%
+\title{SU($n$) ferromagnets near the Heisenberg critical point}%
 \author{Michael A. Perlin}%
@@
 
-SU($n$) symmetries play an important role in contemporary physics.
-Underpinning much of high energy physics, the celebrated SU($n$) gauge
-theory known as Yang-Mills theory is central to our understanding of
-the electroweak and strong forces.  Extensions of Yang-Mills and
-SU($n$) symmetry feature in the most well-studied examples of
-holographic duality\cite{maldacena1999largen} and the connection
-between entanglement and gravity\cite{ryu2006holographic} through the
-anti-de Sitter / conformal field theory (AdS/CFT) correspondence.  In
-a condensed matter setting, SU(2) appears ubiquitously as a symmetry
-of the Hubbard model, with important consequences for the study of
+SU($n$) symmetries play an important role in physics.  Underpinning
+much of high energy physics, the celebrated SU($n$) gauge theory known
+as Yang-Mills theory is central to our understanding of the
+electroweak and strong forces.  Extensions of Yang-Mills and SU($n$)
+symmetry feature in the most well-studied examples of holographic
+duality\cite{maldacena1999largen} and the connection between
+entanglement and gravity\cite{ryu2006holographic} through the anti-de
+Sitter / conformal field theory (AdS/CFT) correspondence.  In a
+condensed matter setting, SU(2) appears ubiquitously as a symmetry of
+the Hubbard model, with important consequences for the study of
 quantum magnetism and high temperature
@@
 thereby enabling the detailed study of these symmetries' consequences
-for fundamental questions in physics, as well as their use in
-practical technological applications.  SU(2)-symmetric interactions,
-for example, can play an important role in the development of quantum
+for fundamental questions in physics, as well as their practical use
+in technological applications.  SU(2)-symmetric interactions, for
+example, can play an important role in the development of quantum
 sensors that make use of entanglement to surpass classical limits on
@@
 SU($n$) symmetries to achieve a technological advantage is still an
-unexplored avenue with enormous potential.
-
-\red{[finish introduction]}
-
-\vfill
-
-\vspace{3cm}
+unexplored avenue of research with enormous potential.
+
+In this work, we generalize some of the techniques used to harness
+SU(2)-symmetric spin interactions for metrological advantage to the
+case of SU($n$).  In doing so, we develop a low-energy effective
+theory of gapped SU($n$) ferromagnets with weakly broken SU($n$)
+symmetry, e.g.~by the presence of external fields or additional
+interaction terms.  The restriction of our theory to ferromagnetic
+interactions can be relaxed to the assumption of
+\begin{enumerate*}
+\item an initial state in the manifold of permutationally-symmetric
+  states, a multilevel generalization of the Dicke manifold for
+  two-level spins, together with the assumption that
+\item this manifold is gapped away from states that break
+  permutational symmetry.
+\end{enumerate*}
+In addition to working out the effective Hamiltonian induced by weak
+SU($n$) symmetry breaking through second order in perturbation theory,
+we develop a diagrammatic method to systematically compute all matrix
+elements of arbitrary spin operators restricted manifolds of weakly
+broken permutation symmetry, thereby enabling the numerical simulation
+of dynamics induced by larger, non-perturbative modifications to the
+SU($n$) Heisenberg model.  While the latter simulations generally
+accumulate uncontrolled truncation errors, we identify criteria under
+which numerical results correctly capture the qualitative dynamical
+behavior of a system.
+
+In the following section, \ref{sec:bare_sun}, we introduce the SU($n$)
+Heisenberg model, finding a convenient way to express its Hamiltonian
+in terms of permutation operators that are agnostic to the value of
+$n$.  In Section \ref{sec:breaking_sun}, we consider the introduction
+of general spin operators, such as inhomogeneous external fields, that
+weakly break SU($n$) symmetry.  We sketch out a derivation of the
+effective Hamiltonian induced by these terms through second order in
+perturbation theory, deferring more detailed calculations to
+Appendices \ref{sec:eigenstates} and \ref{sec:operator_product}.
+Finally, we discuss a numerical method to simulate dynamics within
+manifolds of weak permutational symmetry breaking in Section
+\ref{sec:shell_model}, before concluding in Section
+\ref{sec:conclusions}.  Wherever possible, we will fall back to the
+broadly familiar case of SU(2) in order to facilitate an understanding
+of the general result for SU($n$).
+
+\red{[re-visit / revise the introduction after finishing the remaining
+  text]}
 
@@
 \section{The SU($n$) Heisenberg model}
+\label{sec:bare_sun}
 
@@
 $s_\alpha\equiv\sigma_\alpha/2$, and $\sigma_\alpha$ a single-spin
-Pauli operator.
+Pauli operator.  This permutation operator is equivalently the
+multilevel SWAP operation, which merely swaps the state of two spins.
 
@@
 $n=2$, for example, the PS manifold $\M_0$ is precisely the Dicke
-manifold\cite{dicke1954coherence} of $N$ qubits, spanned by
-${N+2-1\choose2-1}=N+1$ Dicke states $\ket{m_0,m_1}$ for which
-$m_0+m_1=N$ and the integer $m_0$ ($m_1$) indicates the total
-occupation number of qubit state $\ket{0}$ ($\ket{1}$).  The energy of
-any PS state $\ket\psi\in\M_0$ with respect to the interaction
-Hamiltonian $H_0$ in \eqref{eq:H_0} is
+manifold of $N$ qubits, spanned by ${N+2-1\choose2-1}=N+1$ Dicke
+states $\ket{m_0,m_1}$ for which $m_0+m_1=N$ and the integer $m_0$
+($m_1$) indicates the total occupation number of qubit state $\ket{0}$
+($\ket{1}$).  The energy of any PS state $\ket\psi\in\M_0$ with
+respect to the interaction Hamiltonian $H_0$ in \eqref{eq:H_0} is
 \begin{align}
@@
 \section{Breaking SU($n$) symmetry}
+\label{sec:breaking_sun}
 
@@
 
+% briefly add expression for the third order Hamiltonian
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Beyond ground-state perturbation theory}
+\label{sec:shell_model}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{The XXZ model}
+\section{Conclusions}
+\label{sec:conclusions}
 

### 2020-04-04 collective_sun/sun_perturbation.tex
re-wrote the introduction to the Heisenberg model

@@
 \newcommand{\G}{\mathcal{G}}
+\renewcommand{\H}{\mathcal{H}}
 \newcommand{\I}{\mathcal{I}}
@@
 \item an initial state in the manifold of permutationally-symmetric
-  states, a multilevel generalization of the Dicke manifold for
+  states, a multi-level generalization of the Dicke manifold for
   two-level spins, together with the assumption that
@@
 
-% todo: give example of SU(2)
-
-We consider an array of $N$ multilevel spins with SU($n$)-symmetric
-interactions that can be written in the form
-\begin{align}
-  H_0 = \sum_{\set{p,q}\in\C_N\p{2}} h_{pq} \Pi_{pq},
-  &&
-  \Pi_{pq} \equiv \sum_{\mu,\nu\in\ZZ_n} s_{\mu\nu}^{(p)} s_{\nu\mu}^{(q)},
+Before getting to the full SU($n$) Heisenberg model, we first revisit
+the more common case of SU(2).  The SU(2) Heisenberg model consists of
+$N$ two-level spins on a lattice, with spin-aligning interactions
+\begin{align}
+  H_0 = \sum_{\substack{\t{spin pairs}\\\set{i,j}}}
+  h_{ij} \m s_i \c\m s_j,
+  \label{eq:H_SU2}
+\end{align}
+where
+$\m s_i\equiv\p{\sigma_\x^{(i)},\sigma_\y^{(i)},\sigma_\z^{(i)}}/2$ is
+a spin vector defined with Pauli operators $\sigma_\alpha^{(i)}$
+acting on the spin at lattice site $i$.  In the traditional language
+of spin physics, each term of this Hamiltonian essentially imposes an
+energy penalty of $h_{ij}$ to the formation of a spin 0 singlet state
+$\ket{\up\dn}-\ket{\dn\up}$ between spins $i$ and $j$, favoring spin
+triplet states with total spin 1.  Another interpretation of the
+Hamiltonian in \eqref{eq:H_SU2} is illuminated by the observation that
+the two-spin operator $\m s_i\c\m s_j$ is, up to an overall constant,
+essentially the two-qubit SWAP operation $\Pi_{ij}$.  We will also
+refer to $\Pi_{ij}$ as {\it permutation operator}, because its action
+on the Hilbert space $\H\equiv\bigotimes_{k=1}^N\H_k$ of $N$ spins
+with individual Hilbert spaces $\H_k$ is to permute the tensor factors
+of $\H$ associated with spins $i$ and $j$.  Up to identity terms with
+no physical consequence, we then have
+\begin{align}
+  H_0 = \sum_{\substack{\t{spin pairs}\\\set{i,j}}} h_{ij} \Pi_{ij},
+  &&
+  \Pi_{ij} \equiv \sum_{\substack{\t{spin states}\\\mu,\nu}}
+  s_{\mu\nu}^{(i)} s_{\nu\mu}^{(j)}
+  = \sum_{\substack{\t{spin states}\\\mu,\nu}} \op{\mu\nu}{\nu\mu}_{ij},
   \label{eq:H_0}
 \end{align}
-where $p,q$ index individual spins; $\C_N\p{2}$ is the set of all
-subsets (``choices'') of two spins out of $N$; $h_{pq}$ are scalar
-coupling constants; $\mu,\nu$ index states in an orthonormal basis for
-the $n$-dimensional Hilbert space of a single spin; $\ZZ_n$ is the set
-of integers modulo $n$; the operator
-$s_{\mu\nu}^{(p)}\equiv\op{\mu}{\nu}_p$ flips the state of spin $p$ to
-$\ket\nu$ from $\ket\mu$; and the operator $\Pi_{pq}$ permutes the
-states of spins $p$ and $q$.  The permutation operator $\Pi_{pq}$ is
-essentially a multilevel generalization of the SU(2)-symmetric spin
-interaction $\v s_p\c\v s_q$, with $\v s\equiv\p{\v s_\x,s_\y,s_\z}$,
-$s_\alpha\equiv\sigma_\alpha/2$, and $\sigma_\alpha$ a single-spin
-Pauli operator.  This permutation operator is equivalently the
-multilevel SWAP operation, which merely swaps the state of two spins.
-
-If the coefficients $h_{pq}$ of the interaction Hamiltonian $H_0$ are
-all negative, then the ground-state manifold $\M_0$ of the interaction
-Hamiltonian $H_0$ consists of permutationally symmetric (PS) states
-that are simultaneous $+1$ eigenstates of all permutation operators
-$\Pi_{pq}$.  We will assume that all $h_{pq}<0$ throughout these
+% todo: insert visualization of SWAP operation
+where $s_{\mu\nu}^{(i)}\equiv\op{\mu}{\nu}_i$ is a spin-transition
+operator for spin $i$, and $\mu,\nu\in\set{\up,\dn}$ index the states
+of a single spin.  In addition to providing a nice interpretation of
+the Hamiltonian $H_0$ as essentially swapping the states of spins $i$
+and $j$ at a rate $h_{ij}$, the definition of the SU(2)-symmetric
+Heisenberg model in in \eqref{eq:H_0} is straightforward to generalize
+to SU($n$) with arbitrary $n$.  That is, rather than having
+$\mu,\nu\in\set{\up,\dn}$, we simply let
+$\mu,\nu\in\set{1,2,\cdots,n}$ index orthogonal states (sometimes
+called {\it colors} or {\it flavors}) of an $n$-level spin.
+
+If the coefficients $h_{ij}$ of the interaction Hamiltonian $H_0$ are
+all negative, then the ground-state manifold $\M_0$ of $H_0$ consists
+of spin-polarized, {\it permutationally symmetric} (PS) states that
+are simultaneous $+1$ eigenstates of all permutation operators
+$\Pi_{ij}$.  We will assume that all $h_{ij}<0$ throughout these
 notes, but note that this restriction can be relaxed to the assumption
-that the initial state of all spins is permutationally symmetric
-(e.g.~a spin-polarized state).  We will also assume that the PS
-manifold $\M_0$ is gapped away from all orthogonal states by an
-interaction energy $\Delta_{\t{gap}}>0$.  Power-law couplings of the
-form $h_{pq}\sim-1/\abs{p-q}^\alpha$ on a $D$-dimensional lattice, for
-example, always yield a non-vanishing spectral gap $\Delta_{\t{gap}}$
-for long-range interactions with $\alpha\le D$ (see Appendix
-\ref{sec:spectral_gap}).
-
-The PS manifold $\M_0$ is spanned by a basis of states
+that the initial state of all spins is a permutationally-symmetric
+state.  We will also assume that the PS manifold $\M_0$ is gapped away
+from all orthogonal states by an interaction energy
+$\Delta_{\t{gap}}>0$.  Power-law couplings of the form
+$h_{ij}\sim-1/\abs{\m r_i-\m r_j}^\alpha$ on a $D$-dimensional
+lattice, for example, always yield a non-vanishing spectral gap
+$\Delta_{\t{gap}}$ for long-range interactions with $\alpha\le D$ (see
+Appendix \ref{sec:spectral_gap}).  Here $\m r_i$ denotes the position
+of lattice site $i$.
+
+In the case of SU(2), the PS manifold $\M_0$ is precisely the {\it
+  Dicke manifold} spanned by the PS states $\ket{m_\up,m_\dn}$ with a
+definite (integer) number $m_\mu\ge0$ of spins occupying the
+single-spin state $\mu\in\set{\up,\dn}$.  If there are $N$ spins
+total, then Dicke states $\ket{m_\up,m_\dn}$ must have
+$m_\up+m_\dn=N$.  In the case of multi-level spins, the PS manifold
+$\M_0$ is similarly spanned by states
 $\ket{m}=\ket{m_1,m_2,\cdots,m_n}$ with a definite occupation number
-$m_\mu$ of the single-spin state $\mu$, and $\sum_\mu m_\mu=N$.  We
-denote the set of all such assignments of $N$ (identical) spins to $n$
-(distinct) states by $\A_n\p{N}$, such that our standard basis for
-$\M_0$ is $\set{\ket{m}:m\in\A_n\p{N}}$.  The dimension of $\M_0$ is
-then the number of such assignments, which is straightforward to
-determine with a standard ``stars and bars'' calculation:
-$\abs{\A_n\p{N}} = {N+n-1 \choose n-1} \sim N^{n-1}$.  In the case of
-$n=2$, for example, the PS manifold $\M_0$ is precisely the Dicke
-manifold of $N$ qubits, spanned by ${N+2-1\choose2-1}=N+1$ Dicke
-states $\ket{m_0,m_1}$ for which $m_0+m_1=N$ and the integer $m_0$
-($m_1$) indicates the total occupation number of qubit state $\ket{0}$
-($\ket{1}$).  The energy of any PS state $\ket\psi\in\M_0$ with
-respect to the interaction Hamiltonian $H_0$ in \eqref{eq:H_0} is
-\begin{align}
-  E_0 = \sum_{\set{p,q}\in\C_N\p{2}} h_{pq}.
-\end{align}
+$m_\mu\ge0$ of the single-spin state $\mu$, with $\sum_\mu m_\mu=N$.
+These states are essentially labeled by an assignment
+$m=\p{m_1,m_2,\cdots,m_n}$ of $N$ (identical) spins to $n$ (distinct)
+states.  We denote the set of all such assignments by $\A_n\p{N}$,
+such that our standard basis for $\M_0$ is
+$\set{\ket{m}:m\in\A_n\p{N}}$.  The dimension of $\M_0$, or
+equivalently the ground-state degeneracy of $H_0$, is straightforward
+to determine with a standard ``stars and bars'' calculation:
+$\dim\M_0=\abs{\A_n\p{N}} = {N+n-1 \choose n-1} \sim N^{n-1}$.  The
+energy of any PS state $\ket\psi\in\M_0$ with respect to the
+interaction Hamiltonian $H_0$ in \eqref{eq:H_0} is
+\begin{align}
+  E_0 = \sum_{\substack{\t{spin pairs}\\\set{i,j}}} h_{ij}
+  = \sum_{i<j} h_{ij},
+\end{align}
+where $i,j\in\set{1,2,\cdots,N}$ index individual lattice sites.
 
@@
 get
+% todo: define \simeq
 \begin{align}

### 2020-04-09 collective_sun/multilevel_spin_notes.tex
fix typo in multilevel spin notes

@@
 \end{align}
-where $L\in\set{0,1,\cdots,n-1}$ and $M\in\set{-L,-L+1,L}$ index a
-total spin and its projection onto a quantization axis;
+where $L\in\set{0,1,\cdots,n-1}$ and $M\in\set{-L,-L+1,\cdots,L}$
+index a total spin and its projection onto a quantization axis;
 $I\equiv\p{n-1}/2$ is the maximal angular momentum of an $n$-level

### 2020-04-09 collective_sun/sun_perturbation.tex
add sketch to explain permutation operator

@@
 for fundamental questions in physics, as well as their practical use
-in technological applications.  SU(2)-symmetric interactions, for
+in technological applications.  SU(2)-symmetric spin interactions, for
 example, can play an important role in the development of quantum
@@
 Heisenberg model, finding a convenient way to express its Hamiltonian
-in terms of permutation operators that are agnostic to the value of
-$n$.  In Section \ref{sec:breaking_sun}, we consider the introduction
-of general spin operators, such as inhomogeneous external fields, that
+in terms of SWAP operations that are agnostic to the value of $n$.  In
+Section \ref{sec:breaking_sun}, we consider modifying the Hamiltonian
+of the bare Heisenberg model by the addition of spin operators,
+corresponding to e.g.~external fields or additional interactions, that
 weakly break SU($n$) symmetry.  We sketch out a derivation of the
@@
 broadly familiar case of SU(2) in order to facilitate an understanding
-of the general result for SU($n$).
+of the generalizations to SU($n$).
 
@@
 where
-$\m s_i\equiv\p{\sigma_\x^{(i)},\sigma_\y^{(i)},\sigma_\z^{(i)}}/2$ is
-a spin vector defined with Pauli operators $\sigma_\alpha^{(i)}$
-acting on the spin at lattice site $i$.  In the traditional language
-of spin physics, each term of this Hamiltonian essentially imposes an
-energy penalty of $h_{ij}$ to the formation of a spin 0 singlet state
-$\ket{\up\dn}-\ket{\dn\up}$ between spins $i$ and $j$, favoring spin
-triplet states with total spin 1.  Another interpretation of the
-Hamiltonian in \eqref{eq:H_SU2} is illuminated by the observation that
-the two-spin operator $\m s_i\c\m s_j$ is, up to an overall constant,
-essentially the two-qubit SWAP operation $\Pi_{ij}$.  We will also
-refer to $\Pi_{ij}$ as {\it permutation operator}, because its action
-on the Hilbert space $\H\equiv\bigotimes_{k=1}^N\H_k$ of $N$ spins
-with individual Hilbert spaces $\H_k$ is to permute the tensor factors
-of $\H$ associated with spins $i$ and $j$.  Up to identity terms with
-no physical consequence, we then have
+$\m s_i\equiv\f12\p{\sigma_\x^{(i)},\sigma_\y^{(i)},\sigma_\z^{(i)}}$
+is a spin vector defined with Pauli operators $\sigma_\alpha^{(i)}$
+acting on the spin at lattice site $i$.  In the standard language of
+spin physics, each term of this Hamiltonian essentially imposes an
+energy penalty of $-h_{ij}$ to the formation of a spin 0 singlet state
+$\ket{\up\dn}-\ket{\dn\up}$ between spins $i$ and $j$, favoring the
+formation of spin triplet states with total spin 1.  Another
+interpretation of the Hamiltonian in \eqref{eq:H_SU2} is illuminated
+by the observation that the two-spin operator $\m s_i\c\m s_j$ is, up
+to an overall constant, essentially the two-qubit SWAP operation
+$\Pi_{ij}$.  We will also refer to $\Pi_{ij}$ as {\it permutation
+  operator}, because its action on the Hilbert space
+$\H\equiv\bigotimes_{k=1}^N\H_k$ of $N$ spins with individual Hilbert
+spaces $\H_k$ is to permute the tensor factors of $\H$ associated with
+spins $i$ and $j$ (see Figure \ref{fig:sketch_spins}).  Up to identity
+terms with no physical consequence, we then have
 \begin{align}
@@
 \end{align}
-% todo: insert visualization of SWAP operation
 where $s_{\mu\nu}^{(i)}\equiv\op{\mu}{\nu}_i$ is a spin-transition
@@
 to SU($n$) with arbitrary $n$.  That is, rather than having
-$\mu,\nu\in\set{\up,\dn}$, we simply let
-$\mu,\nu\in\set{1,2,\cdots,n}$ index orthogonal states (sometimes
+$\mu,\nu\in\set{\up,\dn}$ index two spin states, we simply let
+$\mu,\nu\in\set{1,2,\cdots,n}$ index the orthogonal states (sometimes
 called {\it colors} or {\it flavors}) of an $n$-level spin.
+
+\begin{figure}
+  \centering
+  \includegraphics[valign=m]{sketch_spins.pdf}
+  $\stackrel{\displaystyle\Pi_{ij}}
+  {\displaystyle\longleftrightarrow}$
+  \includegraphics[valign=m]{sketch_spins_swap.pdf}
+  \caption{The permutation operator $\Pi_{ij}$ swaps the states of
+    spins $i$ and $j$, which in the case of SU(2) can be represented
+    by magnetization vectors in 3-dimensional space.  In the case of
+    SU($n$), these vectors live in a higher-dimensional space without
+    a clear geometric interpretation, but the interpretation of the
+    permutation operator $\Pi_{ij}$ remains the same.}
+  \label{fig:sketch_spins}
+\end{figure}
 
@@
 \label{sec:breaking_sun}
+
+\vspace{3cm}
 

### 2020-04-10 collective_sun/sun_perturbation.tex
added mention of valence bond solids to SU(n) ferromagnet paper other minor text changes

@@
 Hubbard model to SU($n$) has led to predictions of exotic phases of
-matter such as chiral spin liquids\cite{hermele2009mott,
-  hermele2011topological, chen2016syntheticgaugefield,
-  nataf2016chiral}, as well as the potential to perform universal
-topological quantum computation\cite{freedman2004class,
-  nayak2008nonabelian}.  Furthermore, the consideration of disordered
-SU($n$) spin models has opened analytically tractable avenues for
-studying quantum chaos and information
-scrambling\cite{sachdev1993gapless, bentsen2019integrable}.
+matter such as valence bond solids\cite{read1989valencebond,
+  rokhsar1990quadratic, kaul2012lattice, hermele2011topological} and
+chiral spin liquids\cite{hermele2009mott, hermele2011topological,
+  chen2016syntheticgaugefield, nataf2016chiral}, as well as the
+potential to perform universal topological quantum
+computation\cite{freedman2004class, nayak2008nonabelian}.
+Furthermore, the consideration of disordered SU($n$) spin models has
+opened analytically tractable avenues for studying quantum chaos and
+information scrambling\cite{sachdev1993gapless,
+  bentsen2019integrable}.
 
@@
 broken permutation symmetry, thereby enabling the numerical simulation
-of dynamics induced by larger, non-perturbative modifications to the
-SU($n$) Heisenberg model.  While the latter simulations generally
-accumulate uncontrolled truncation errors, we identify criteria under
-which numerical results correctly capture the qualitative dynamical
-behavior of a system.
+of dynamics induced by non-perturbative modifications to the SU($n$)
+Heisenberg model.  While the latter simulations generally accumulate
+uncontrolled truncation errors, we identify criteria under which
+numerical results correctly capture the qualitative dynamical behavior
+of a system.
 
@@
 to an overall constant, essentially the two-qubit SWAP operation
-$\Pi_{ij}$.  We will also refer to $\Pi_{ij}$ as {\it permutation
+$\Pi_{ij}$.  We will refer to $\Pi_{ij}$ as {\it permutation
   operator}, because its action on the Hilbert space
@@
 that the initial state of all spins is a permutationally-symmetric
-state.  We will also assume that the PS manifold $\M_0$ is gapped away
-from all orthogonal states by an interaction energy
-$\Delta_{\t{gap}}>0$.  Power-law couplings of the form
-$h_{ij}\sim-1/\abs{\m r_i-\m r_j}^\alpha$ on a $D$-dimensional
+state (e.g.~a spin-polarized state).  We will also assume that the PS
+manifold $\M_0$ is gapped away from all orthogonal states by an
+interaction energy $\Delta_{\t{gap}}>0$.  Power-law couplings of the
+form $h_{ij}\sim-1/\abs{\m r_i-\m r_j}^\alpha$ on a $D$-dimensional
 lattice, for example, always yield a non-vanishing spectral gap
@@
 to determine with a standard ``stars and bars'' calculation:
-$\dim\M_0=\abs{\A_n\p{N}} = {N+n-1 \choose n-1} \sim N^{n-1}$.  The
-energy of any PS state $\ket\psi\in\M_0$ with respect to the
+$\dim\p{\M_0}=\abs{\A_n\p{N}} = {N+n-1 \choose n-1} \sim N^{n-1}$.
+The energy of any PS state $\ket\psi\in\M_0$ with respect to the
 interaction Hamiltonian $H_0$ in \eqref{eq:H_0} is
@@
 
-% give examples of these manifolds in SU(2)
+% todo: give examples of these manifolds in SU(2)
+% todo: clarify connection between multi-body excitation manifolds
+%       and manifolds of fixed total spin in the SU(2) case
 

### 2020-04-25 collective_sun/sun_perturbation.tex
overhauled lots of conventions in the SU(n) perturbation notes and shell model codes simplified the section on breaking SU(n) symmetry (perturbatively)

@@
 \renewcommand{\sp}[1]{\left[#1\right]} % square parenthesis
+\newcommand{\spp}[1]{\sp{\sp{#1}}} % double square parenthesis
 \renewcommand{\set}[1]{\left\{#1\right\}} % curly parenthesis
@@
 \renewcommand{\c}{\cdot} % inner product
-
 \newcommand{\m}{\bm} % bold symbol
@@
 
+\renewcommand{\i}{\mathrm{i}\mkern1mu}
 \renewcommand{\d}{\text{d}}
@@
 \renewcommand{\P}{\mathcal{P}}
+\newcommand{\Q}{\mathcal{Q}}
 \renewcommand{\S}{\mathcal{S}}
@@
 \newcommand{\EQPS}{=_{\text{PS}}}
-\newcommand{\col}{\underline}
 
@@
 \DeclareMathOperator{\diag}{diag}
-\newcommand{\ul}{\underline}
 
@@
 \newcommand{\ut}{\undertilde}
+\newcommand{\ot}{\widetilde}
+\newcommand{\col}{\underline}
+\newcommand{\mean}{\overline}
 
@@
 \begin{align}
-  H_0 = \sum_{\substack{\t{spin pairs}\\\set{i,j}}}
-  h_{ij} \m s_i \c\m s_j,
+  H_0 = \sum_{i\ne j} h_{ij} \m s_i \c\m s_j,
   \label{eq:H_SU2}
 \end{align}
-where
-$\m s_i\equiv\f12\p{\sigma_\x^{(i)},\sigma_\y^{(i)},\sigma_\z^{(i)}}$
-is a spin vector defined with Pauli operators $\sigma_\alpha^{(i)}$
-acting on the spin at lattice site $i$.  In the standard language of
-spin physics, each term of this Hamiltonian essentially imposes an
-energy penalty of $-h_{ij}$ to the formation of a spin 0 singlet state
-$\ket{\up\dn}-\ket{\dn\up}$ between spins $i$ and $j$, favoring the
-formation of spin triplet states with total spin 1.  Another
-interpretation of the Hamiltonian in \eqref{eq:H_SU2} is illuminated
-by the observation that the two-spin operator $\m s_i\c\m s_j$ is, up
-to an overall constant, essentially the two-qubit SWAP operation
-$\Pi_{ij}$.  We will refer to $\Pi_{ij}$ as {\it permutation
-  operator}, because its action on the Hilbert space
+where $i,j$ index individual spins; $h_{ij}$ are scalar coefficients
+with $h_{ij}=h_{ji}$ and $h_{ii}=0$; and
+$\m s_i\equiv\p{s_\x^{(i)},s_\y^{(i)},s_\z^{(i)}}$ is a vector of spin
+operators $s_\alpha^{(i)}\equiv\sigma_\alpha^{(i)}/2$ defined by Pauli
+operators $\sigma_\alpha^{(i)}$ acting on spin $i$.  In the standard
+language of spin physics, each term of this Hamiltonian essentially
+imposes an energy penalty of $-2h_{ij}$ to the formation of a spin 0
+singlet state $\ket{\up\dn}-\ket{\dn\up}$ between spins $i$ and $j$,
+favoring the formation of spin triplet states with total spin 1.
+Another interpretation of the Hamiltonian in \eqref{eq:H_SU2} is
+illuminated by the observation that the two-spin operator
+$\m s_i\c\m s_j$ is, up to an overall constant, essentially the
+two-qubit SWAP operation $\Pi_{ij}$.  We will refer to $\Pi_{ij}$ as
+{\it permutation operator}, because its action on the Hilbert space
 $\H\equiv\bigotimes_{k=1}^N\H_k$ of $N$ spins with individual Hilbert
@@
 \begin{align}
-  H_0 = \sum_{\substack{\t{spin pairs}\\\set{i,j}}} h_{ij} \Pi_{ij},
-  &&
-  \Pi_{ij} \equiv \sum_{\substack{\t{spin states}\\\mu,\nu}}
-  s_{\mu\nu}^{(i)} s_{\nu\mu}^{(j)}
-  = \sum_{\substack{\t{spin states}\\\mu,\nu}} \op{\mu\nu}{\nu\mu}_{ij},
+  H_0 = \sum_{i<j} h_{ij} \Pi_{ij},
+  &&
+  \Pi_{ij} \equiv \sum_{\mu,\nu} s_{\mu\nu}^{(i)} s_{\nu\mu}^{(j)}
+  = \sum_{\mu,\nu} \op{\mu\nu}{\nu\mu}_{ij}
   \label{eq:H_0}
@@
 $\mu,\nu\in\set{\up,\dn}$ index two spin states, we simply let
-$\mu,\nu\in\set{1,2,\cdots,n}$ index the orthogonal states (sometimes
-called {\it colors} or {\it flavors}) of an $n$-level spin.
+$\mu,\nu\in\set{1,2,\cdots,n}$ index the states (sometimes called {\it
+  colors} or {\it flavors}) of an $n$-level spin.
 
@@
 interaction energy $\Delta_{\t{gap}}>0$.  Power-law couplings of the
-form $h_{ij}\sim-1/\abs{\m r_i-\m r_j}^\alpha$ on a $D$-dimensional
-lattice, for example, always yield a non-vanishing spectral gap
-$\Delta_{\t{gap}}$ for long-range interactions with $\alpha\le D$ (see
-Appendix \ref{sec:spectral_gap}).  Here $\m r_i$ denotes the position
-of lattice site $i$.
+form $h_{ij}\sim-1/\abs{i-j}^\alpha$ on a $D$-dimensional lattice, for
+example, always yield a non-vanishing spectral gap $\Delta_{\t{gap}}$
+for long-range interactions with $\alpha\le D$ (see Appendix
+\ref{sec:spectral_gap}).
 
@@
 $\ket{m}=\ket{m_1,m_2,\cdots,m_n}$ with a definite occupation number
-$m_\mu\ge0$ of the single-spin state $\mu$, with $\sum_\mu m_\mu=N$.
+$m_\mu\ge0$ of the single-spin state $\mu$, and $\sum_\mu m_\mu=N$.
 These states are essentially labeled by an assignment
@@
 \begin{align}
-  E_0 = \sum_{\substack{\t{spin pairs}\\\set{i,j}}} h_{ij}
-  = \sum_{i<j} h_{ij},
-\end{align}
-where $i,j\in\set{1,2,\cdots,N}$ index individual lattice sites.
+  E_0 = \sum_{i<j} h_{ij} = \f12 \sum_{i\ne j} h_{ij}.
+\end{align}
 
@@
 
-\vspace{3cm}
-
-% todo: simplify, give examples at every step
-% todo: consider special case before a general one
-% todo: new symbol for X_w, e.g. O(X,w)
-
-In order to induce non-trivial dynamics for initially PS states
-$\ket\psi\in\M_0$, which are otherwise stationary eigenstates of the
-interaction Hamiltonian $H_0$ in \eqref{eq:H_0}, we will add to the
-Hamiltonian an operator $\hat\O$ of the general form
-\begin{align}
-  \hat\O \equiv \sum_{\p{X,\m w}\in\O} X_{\m w},
-  &&
-  X_{\m w} \equiv \f1{M!} \sum_{k\in\D_N\p{M}} w_k X_k,
-  \label{eq:pert}
-\end{align}
-where $\hat\O$ is constructed from the set $\O\equiv\set{\p{X,\m w}}$
-of $M$-spin operators $X$ together with dimension-$M$ (i.e.~$M$-index)
-tensors $\m w$; $k=\p{k_1,k_2,\cdots,k_M}$ is a list of of $M$ spins
-that $X_k$ acts on; $w_k$ is a scalar entry of $\m w$; and
+A common task, for example in state preparation protocols or quenched
+dynamics experiments, is to first initialize in some trivial uniform
+product state, $\ket\phi^{\otimes N}$, and then to let it evolve under
+a (possibly time-dependent) Hamiltonian that induces non-trivial
+dynamics.  The permutational symmetry of states such as
+$\ket\phi^{\otimes N}\in\M_0$, however, makes them eigenstates of the
+SU($n$)-symmetric interaction Hamiltonian $H_0$ in \eqref{eq:H_0}.  In
+order to induce non-trivial dynamics, we therefore need to add new
+terms to the Hamiltonian that break its SU($n$) symmetry.
+
+As a concrete example, this sort of scenario can be used in the SU(2)
+case to generate spin squeezed states that harness entanglement to
+surpass classical limits on measurement
+precision\cite{he2019engineering}\red{[cite XXZ squeezing paper]}.
+Starting with a general (gapped) SU(2)-symmetric Hamiltonian of the
+form in \eqref{eq:H_SU2} we can consider the addition of Ising
+interactions
+\begin{align}
+  H_{\t{Ising}} = \sum_{i\ne j} J_{ij} s_\z^{(i)} s_\z^{(j)}.
+  \label{eq:Ising}
+\end{align}
+If these interactions are sufficiently weak for perturbative
+treatment, as in the case of the XXZ model near the Heisenberg
+critical point, then the effective Hamiltonian induced by
+$H_{\t{Ising}}$ on initially PS states $\ket\psi\in\M_0$ is the
+one-axis twisting (OAT) Hamiltonian\cite{ma2011quantum}
+\begin{align}
+  H_{\t{OAT}} = \chi S_\z^2,
+  \label{eq:OAT}
+\end{align}
+where the collective operator $S_\z\equiv\sum_{j=1}^Ns_\z^{(j)}$, and
+the squeezing strength $\chi$ is equal to the mean Ising coupling
+strength $J_{ij}$.  If Ising interactions are unavailable, the same
+effect can be achieved by applying an inhomogeneous magnetic field,
+\begin{align}
+  H_{\t{B}} = \sum_i B_i s_\z^{(i)},
+  \label{eq:magnetic}
+\end{align}
+which gives rise to an OAT Hamiltonian at second order in perturbation
+theory, with a squeezing strength $\chi$ that is proportional to the
+variance of the magnetic field strength $B_i$.
+
+In order to go beyond first-order treatment of two-body interactions
+such as \eqref{eq:Ising}, or beyond second order in the treatment of
+single-body terms such as \eqref{eq:magnetic}, it will turn out to be
+necessary to introduce $M$-body operators with $M>2$.  Generalizing
+the above examples to the case of SU($n$) systems, we therefore
+consider a general $M$-body operator of the form
+\begin{align}
+  \O\p{\m w,X} \equiv \sum_{k\in\D_N\p{M}} w_k X_k,
+  \label{eq:multi_body_op}
+\end{align}
+where $\m w$ is a dimension-$M$ (i.e.~$M$-index) tensor of scalar
+coefficients $w_k\equiv w_{k_1k_2\cdots k_M}$; $X$ is an $M$-spin
+operator, e.g.~$s_\z\otimes s_\z$ in the case of Ising interactions
+with $M=2$; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is a list of the
+individual spins $k_i\in\ZZ_N$ that the operator
+$X_k\equiv X_{k_1k_2\cdots k_M}$ acts on; and
 \begin{align}
   \D_N\p{M} \equiv
-  \set{ k \in \ZZ_N^M: \t{all entries of}~k~\t{are distinct} },
+  \set{ k \in \ZZ_N^M : \t{all entries $k_i$ of $k$ are distinct } },
   \label{eq:off_diags}
 \end{align}
-is the strictly off-diagonal part of $\ZZ_N^M$.  The factor of $1/M!$
-is included in \eqref{eq:pert} by convention to average over the $M!$
-terms addressing any given set of spins.  To simplify language and
-notation, we will assume throughout this work that $\hat\O$ is
-homogeneous in the sense that $M$ is the same for all
-$\p{X,\m w}\in\O$.  A generalization to the case of inhomogeneous
-operators $\hat\O$ is straightforward.
-
-Single-body ($M=1$) operators correspond to classical external fields,
-e.g.~$s_\z$ for a magnetic field, and two-body ($M=2$) operators
-correspond to two-body interactions that deviate from $H_0$, such as
-$s_\z\otimes s_\z$ interactions in an XXZ model near the Heisenberg
-critical point.  While we would be hard-pressed to find a physical
-implementation of multi-body spin interactions with $M>2$, considering
-this case will be necessary to construct low-lying eigenstates of
-$H_0$, which will in turn enable simulating the dynamics induced by
-non-perturbative additions $\hat\O$ to the SU($n$) Heisenberg model,
-with operator norms $\norm*{\hat\O}$ comparable to the spectral gap
-$\Delta_{\t{gap}}$ of $H_0$.
-
-When the operator norm of $\hat\O$ is sufficiently small, namely
-$\norm*{\hat\O}<\Delta_{\t{gap}}/2$, we can treat its effect on the PS
-manifold $\M_0$ perturbatively.  The effective Hamiltonian
-$H_{\t{eff}}$ induced by $\hat\O$ on the ground-state manifold $\M_0$
-through second order in perturbation theory
-is\cite{bravyi2011schrieffer, perlin2019effective}
-\begin{align}
-  H_{\t{eff}} = H_{\t{eff}}^{(1)} + H_{\t{eff}}^{(2)},
-  &&
-  H_{\t{eff}}^{(1)} = \P_0 \hat\O \P_0,
+is the strictly ``off-diagonal'' part of $\ZZ_N^M$, which is necessary
+to identify for a consistent definition of $X_k$ as an $M$-body
+operator.  In this notation, we can write the magnetic field and
+Hamiltonians in \eqref{eq:magnetic} and \eqref{eq:Ising} as
+$H_{\t{B}}=\O\p{\m B,s_\z}$ and
+$H_{\t{Ising}}=\O\p{\m J,s_\z\otimes s_\z}$.
+
+If the addition $\O\p{\m w,X}$ to the SU($n$)-symmetric Hamiltonian
+$H_0$ in \eqref{eq:H_0} is sufficiently small, then we can treat the
+effect of $\O\p{\m w,X}$ on the PS manifold $\M_0$ perturbatively.
+The effective Hamiltonians $H_{\t{eff}}^{(1)}$ and $H_{\t{eff}}^{(2)}$
+induced by $\O\p{\m w,X}$ on $\M_0$ at first and second order in
+perturbation theory are\cite{bravyi2011schrieffer}
+\begin{align}
+  H_{\t{eff}}^{(1)} = \P_0 \O\p{\m w,X} \P_0,
   &&
   H_{\t{eff}}^{(2)} = - \sum_{\Delta\ne0}
-  \f1\Delta \P_0 \hat\O \P_\Delta \hat\O \P_0,
+  \f1\Delta \P_0 \O\p{\m w,X} \P_\Delta \O\p{\m w,X} \P_0,
   \label{eq:H_eff}
@@
 projector onto $\M_0$ itself, and $\E_0=\M_0$.  The first order
-effective Hamiltonian $H_{\t{eff}}^{(1)}$ simply projects the
-perturbation $\hat\O$ onto the permutationally-symmetric manifold
-$\M_0$, and takes the form
-\begin{align}
-  H_{\t{eff}}^{(1)} \EQPS \sum_{\p{X,\m w}\in\O} \bar w\,\col{X},
+effective Hamiltonian $H_{\t{eff}}^{(1)}$ simply projects
+$\O\p{\m w,X}$ onto the PS manifold $\M_0$, and takes the form
+\begin{align}
+  H_{\t{eff}}^{(1)} \EQPS \mean w \, \col X,
   \label{eq:H_eff_1}
@@
 where $\EQPS$ denotes equality under a restriction to $\M_0$; the
-coefficient $\bar w$ is the average of all $w_k$; and the collective
-operator $\col{X}$ is a sum of all $X_k$ averaged over permutations of
-any particular choice of spins $k\in\D_N\p{M}$.  To write out the
-collective operator $\col{X}$ explicitly, we define the set \begin{align}
-  \C_N\p{M} \equiv \set{ k \subset \ZZ_N: \abs{k} = M },
-  \label{eq:choices}
-\end{align}
-is the set of all subsets (``choices'') of $M$ elements from $\ZZ_N$,
-and the set $\pi\p{M}$ of all permutations of $M$ elements, in terms
-of which
-\begin{align}
-  \bar w \equiv \f1{\abs{\D_N\p{M}}}
+coefficient $\mean w$ is the average of all $w_k$; and $\col X$ is a
+collective version of $X$:
+\begin{align}
+  \mean w \equiv \f1{\abs{\D_N\p{M}}}
   \sum_{k\in\D_N\p{M}} w_k,
   &&
-  \col{X} \equiv \f1{\abs{\pi\p{M}}} \sum_{\sigma\in\pi\p{M}}
-  \sum_{k\in\C_N\p{M}} X_{\sigma\p{k}},
-\end{align}
-where $\abs{\D_N\p{M}}=\prod_{j=0}^{M-1}\p{N-j}$ and
-$\abs{\pi\p{M}}=M!$.  Here $k\in\C_N\p{M}$ defines which spins
-$X_{\sigma\p{k}}$ acts on, while $\sigma\in\pi\p{M}$ determines which
-tensor factor of $X$ each of these spins is associated with.  An
-immediate consequence of \eqref{eq:H_eff_1} is the fact that
-perturbing SU(2)-symmetric interactions by two-body $s_\z\otimes s_\z$
-interactions yields the one-axis twisting (OAT)
-Hamiltonian\cite{kitagawa1993squeezed, ma2011quantum}
-$H_{\t{OAT}}=\chi S_\z^2$ at first order in perturbation theory, with
-the OAT strength $\chi$ equal to the mean coefficient of the two-body
-$s_\z\otimes s_\z$ terms.
-
-The second order effective Hamiltonian $H_{\t{eff}}^{(2)}$ is not as
-straightforward to simplify due to the presence of a projector
-$\P_\Delta$ onto the manifold $\E_\Delta$ of states with excitation
-energy $\Delta$.  For each term $X_{\m w}$ of $\hat\O$, this projector
-essentially picks off the part of $X_{\m w}$ that is strictly
-off-diagonal with respect to the ground- and excited-state manifolds
-$\E_0$ and $\E_\Delta$.  We therefore need to decompose $X_{\m w}$
-into components that generate states of definite excitation energy
-when acting on PS states $\ket\psi\in\M_0$.  The SU($n$) symmetry of
-$H_0$ enables such a decomposition of the form
-\begin{align}
-  H_0 X_{\m w}\ket\psi
-  = \sum_\Delta \p{E_0+\Delta} X_{\m w_\Delta} \ket\psi,
+  \col X \equiv \sum_{k\in\D_N\p{M}} X_k,
+\end{align}
+with $\abs{\D_N\p{M}}=\prod_{j=0}^{M-1}\p{N-j}$.  In the case of a
+magnetic field $s_\z$ or Ising interactions $s_\z\otimes s_\z$, for
+example,
+\begin{align}
+  \col{s_\z} = \sum_i s_\z^{(i)} = S_\z,
+  &&
+  \col{s_\z\otimes s_\z}
+  = \sum_{i\ne j} s_\z^{(i)} s_\z^{(j)}
+  = S_\z^2 - \f14 N.
+\end{align}
+The second order effective Hamiltonian $H_{\t{eff}}^{(2)}$ in
+\eqref{eq:H_eff} takes more work to simplify due to the presence of a
+projector $\P_\Delta$ onto the manifold $\E_\Delta$ of states with
+excitation energy $\Delta$.  This projector essentially picks off the
+part of $\O\p{\m w,X}$ that is strictly off-diagonal with respect to
+the ground- and excited-state manifolds $\E_0$ and $\E_\Delta$.  We
+therefore need to decompose $\O\p{\m w,X}$ into components that
+generate states of definite excitation energy when acting on PS states
+$\ket\psi\in\M_0$.  The SU($n$) symmetry of $H_0$ enables such a
+decomposition of the form
+\begin{align}
+  H_0 \O\p{\m w,X} \ket\psi
+  = \sum_\Delta \p{E_0+\Delta} \O\p{\m w_\Delta,X} \ket\psi,
 \end{align}
@@
 $\hat{\m h}$ is determined entirely by the couplings couplings
-$h_{pq}$ of $H_0$.  While the computational cost of diagonalizing
+$h_{ij}$ of $H_0$.  While the computational cost of diagonalizing
 $\hat{\m h}$ is nominally $O\p{N^{3M}}$, spatial symmetries of $H_0$
-and $X_{\m w}$ such as translational invariance or isotropy can
-drastically reduce the complexity of this eigenvalue problem.  We
-discuss this reduction in Appendix \ref{sec:symmetries}.
-
-Equipped with the decomposition of $X_{\m w}$ into terms that generate
-states of definite excitation energy, we can expand
+and $\O\p{\m w,X}$ such as translational invariance or isotropy can
+drastically reduce the complexity of this eigenvalue problem, as we
+discuss in Appendix \ref{sec:symmetries}.
+
+Equipped with the decomposition of $\O\p{\m w,X}$ into terms
+$\O\p{\m w_\Delta,X}$ that generate states of definite excitation
+energy $\Delta$, we can expand
 \begin{align}
   H_{\t{eff}}^{(2)} = -\sum_{\Delta\ne0} \f1\Delta
-  \sum_{\p{X,\m v},\p{Y,\m w}\in\O}
-  \P_0 X_{\m v_\Delta} Y_{\m w_\Delta} \P_0.
-\end{align}
-Simplifying product $\P_0 X_{\m v_\Delta} Y_{\m w_\Delta} \P_0$
-generally yields $r$-body operators for all positive integers
-$r\le2M$.  We provide a general prescription for simplifying such
-products in Appendix \ref{sec:operator_product}, and explicitly work
-out single-body ($M=1$) and two-body ($M=2$) cases in Appendices
+  \P_0 \O\p{\m w_\Delta,X}^2 \P_0.
+\end{align}
+Simplifying the product $\P_0 \O\p{\m w_\Delta,X}^2 \P_0$ generally
+yields $r$-body operators for all positive integers $r\le2M$.  We
+provide a general prescription for simplifying such products in
+Appendix \ref{sec:operator_product}, and explicitly work out the
+single-body ($M=1$) and two-body ($M=2$) cases in Appendices
 \ref{sec:single_pair_prod} and \ref{sec:two_pair_prod}.  If $X$ is a
-single-body operator ($M=1$), then
+single-body operator, then
 \begin{align}
   H_{\t{eff}}^{(2)}
-  = \sum_{\Delta\ne0} \sum_{\p{X,\m v},\p{Y,\m w}\in\O}
-  \f{\cov\p{\m v_\Delta,\m w_\Delta}}{\p{N-1}\Delta}
-  \p{\col{X}\,\col{Y} - N \col{XY}},
+  = \sum_{\Delta\ne0} \f{\var\p{\m w_\Delta}}{\p{N-1}\Delta}
+  \p{\col{X}^2 - N \col{X^2}},
   \label{eq:H_eff_2_1}
@@
 \begin{align}
-  \cov\p{\m v,\m w} \equiv
-  \f1N \sum_{p\in\ZZ_N} \p{v_p-\bar v}\p{w_p-\bar w}
-\end{align}
-is the covariance between $\m v$ and $\m w$.  In the SU(2) case,
-\eqref{eq:H_eff_2_1} implies that an inhomogeneous magnetic field
-($X\propto s_\z$) yields an OAT Hamiltonian at second order
+  \var\p{\m w} \equiv \f1{\abs{\D_N\p{M}}}
+  \sum_{k\in\D_N\p{M}} \p{w_k - \mean w}^2
+\end{align}
+is the variance of $\m w$.  In the SU(2) case, this result implies
+that an inhomogeneous magnetic field ($X=s_\z$) yields the OAT
+Hamiltonian $H_{\t{OAT}}$ in \eqref{eq:OAT} at second order
 perturbation theory.
 
-% briefly add expression for the third order Hamiltonian
+\newpage
 
@@
 \label{sec:shell_model}
+
+\red{[note: this section is not yet finished or simplified]}
 
@@
 We then denote the a direct sum of all $\M_r$ for $r\le M$ by
-$\ul\M_M$, i.e.~$\ul\M_M\equiv\bigoplus_{r\le M}\M_r$.  The manifold
-$\ul\M_M$ consists of all states accessible by the action of $r$-body
+$\col\M_M$, i.e.~$\col\M_M\equiv\bigoplus_{r\le M}\M_r$.  The manifold
+$\col\M_M$ consists of all states accessible by the action of $r$-body
 operators with $r\le M$ on PS states in $\M_0$.  These definitions are
@@
 By construction, the $M$-body operator $X_{\m w}$ generally couples PS
-states in $\M_0$ to asymmetric states in $\ul\M_M$.  If the coupling
+states in $\M_0$ to asymmetric states in $\col\M_M$.  If the coupling
 between $\M_0$ and its complement is weak compared to the spectral gap
@@
 $\norm{X_{\m w}}\ge\Delta_{\t{gap}}/2$, in which case the operator
-$X_{\m w}$ first couples $\M_0$ to $\ul\M_M$, then $\ul\M_M$ to
-$\ul\M_{2M}$, and so on.  Even if
+$X_{\m w}$ first couples $\M_0$ to $\col\M_M$, then $\col\M_M$ to
+$\col\M_{2M}$, and so on.  Even if
 $\norm{X_{\m w}}>\Delta_{\t{gap}}/2$, however, the energetic penalty
@@
 the dynamics of an initial PS state $\ket\psi\in\M_0$ by projecting
-$X_{\m w}$ onto $\ul\M_{r_{\t{max}}}$ for some $r_{\t{max}}\ge M$,
+$X_{\m w}$ onto $\col\M_{r_{\t{max}}}$ for some $r_{\t{max}}\ge M$,
 resulting in a theory that resembles first-order perturbation theory
-on $\ul\M_{r_{\t{max}}}$.  In the perturbative regime
+on $\col\M_{r_{\t{max}}}$.  In the perturbative regime
 $\norm{X_{\m w}}<\Delta_{\t{gap}}/2$, such a theory captures all
@@
 truncating the Hilbert space of a spin system past
-$\ul\M_{r_{\t{max}}}$ is not clear a priori.  Nonetheless, simulations
-restricted to $\ul\M_{r_{\t{max}}}$ can be benchmarked by examining
-the maximum populations within all $\M_r$ for $r\le r_{\t{max}}$
-throughout simulation.  As long as
+$\col\M_{r_{\t{max}}}$ is not clear a priori.  Nonetheless,
+simulations restricted to $\col\M_{r_{\t{max}}}$ can be benchmarked by
+examining the maximum populations within all $\M_r$ for
+$r\le r_{\t{max}}$ throughout simulation.  As long as
 \begin{enumerate*}
@@
 
-In order to project an operator $X_{\m w}$ onto $\ul\M_{r_{\t{max}}}$,
-we first need to construct a basis for $\ul\M_{r_{\t{max}}}$, and then
-compute all matrix elements of $X_{\m w}$ in this basis.  Our strategy
-will be to first identify a basis for $\M_0$, and then find, for each
-$r=1,\cdots,r_{\t{max}}$, a suitable set of $r$-body operators
-$\set{O}$ and dimension-$r$ tensors $\set{\m u}$ for which the states
+In order to project an operator $X_{\m w}$ onto
+$\col\M_{r_{\t{max}}}$, we first need to construct a basis for
+$\col\M_{r_{\t{max}}}$, and then compute all matrix elements of
+$X_{\m w}$ in this basis.  Our strategy will be to first identify a
+basis for $\M_0$, and then find, for each $r=1,\cdots,r_{\t{max}}$, a
+suitable set of $r$-body operators $\set{O}$ and dimension-$r$ tensors
+$\set{\m u}$ for which the states
 $\ket{O,\m u,m}\propto O_{\m u}\ket{m}$ form an orthonormal basis for
 $\M_r$.  Matrix elements of $X_{\m w}$ with respect to this basis for
-$\ul\M_{r_{\t{max}}}$ then take the form
+$\col\M_{r_{\t{max}}}$ then take the form
 \begin{align}
@@
 
+\red{[look at \cite{rai1988correlated} for classifying asymmetric
+  states]}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Here we show that power-law SU($n$)-symmetric interactions of the form
-in \eqref{eq:H_0} with $h_{pq}=-h/\abs{p-q}^\alpha$ yield a
+in \eqref{eq:H_0} with $h_{ij}=-h/\abs{i-j}^\alpha$ yield a
 non-vanishing many-body interaction energy gap when $\alpha\le D$,
@@
 get
-% todo: define \simeq
+% todo: define \simeq ?
 \begin{align}
@@
 \begin{align}
-  \Delta = h \epsilon^{\alpha-D} \I_{D\epsilon},
-  &&
-  \I_{D\epsilon}
+  \Delta = h \epsilon^{\alpha-D} \I_D\p{\epsilon},
+  &&
+  \I_D\p{\epsilon}
   \equiv \int_{\TT_1^D\setminus\TT_\epsilon^D} \d^Dx\,
@@
 \end{align}
-where the integral $\I_{D\epsilon}$ is defined using the interval
-$\TT_a\equiv\p{-a,a}$.  The integrand of $\I_{D\epsilon}$ is strictly
-positive and well-behaved on the entirety of its domain except for the
-origin, where depending on the value of $\alpha$ the integrand may
-vanish or diverge as $\abs{x}\to0$.  Together, these facts mean that
-\begin{align}
-  \I_{D\epsilon} \stackrel{\epsilon\to0}{\sim} \epsilon^{-\gamma},
+where the integral $\I_D\p{\epsilon}$ is defined using the interval
+$\TT_a\equiv\p{-a,a}$.  The integrand of $\I_D\p{\epsilon}$ is
+strictly positive and well-behaved on the entirety of its domain
+except for the origin, where depending on the value of $\alpha$ the
+integrand may vanish or diverge as $\abs{x}\to0$.  Together, these
+facts mean that
+\begin{align}
+  \I_D\p{\epsilon} \stackrel{\epsilon\to0}{\sim} \epsilon^{-\gamma},
   &&
@@
 this behavior, we note that the asymptotic dependence of the integral
-$\I_{D\epsilon}$ on $\epsilon$ is determined by the behavior of its
+$\I_D\p{\epsilon}$ on $\epsilon$ is determined by the behavior of its
 integrand when $\abs{x}\sim\epsilon$, in which case
@@
 \begin{align}
-  \I_{D\epsilon}
+  \I_D\p{\epsilon}
   \sim \int_{\TT_1^D\setminus\TT_\epsilon^D} \d^Dx\,
@@
 \begin{align}
-  \I_{D\epsilon} \lesssim
+  \I_D\p{\epsilon} \lesssim
   \int_{\TT_1^D\setminus\TT_\epsilon^D} \d^Dx\,
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Constructing excitation energy eigenstates}
+\section{Generating excitation energy eigenstates}
 \label{sec:eigenstates}
 
-In this section, we simplify the products of the form
-$H_0X_{\m w}\ket\psi$ to find conditions under which a multi-body
-perturbation $X_{\m w}$ generates eigenstates of the SU($n$)-symmetric
-interaction Hamiltonian $H_0$ in \eqref{eq:H_0} when acting on a
-permutationally symmetric state $\ket\psi\in\M_0$.  Here $X_{\m w}$ is
-an $M$-body perturbation defined by
-\begin{align}
-  X_{\m w}
-  \equiv \f1{M!} \sum_{k\in\D_N\p{M}} w_k X_k,
-\end{align}
-where $X$ is an $M$-spin operator; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is
-a list of $M$ spins that $X_k$ acts on; $\m w$ is a dimension-$M$
-(i.e.~$M$-index) tensor with scalar entries $w_k$; and $\D_N\p{M}$ is
-the strictly off-diagonal part of $\ZZ_N^M$, defined in
-\eqref{eq:off_diags}.
+Both perturbation theory and beyond-perturbative simulations of the
+dynamics induced by an operator $\O\p{\m w,X}$ require us to decompose
+$\O\p{\m w,X}$ into a sum of terms that generate eigenstates of the
+SU($n$)-symmetric interaction Hamiltonian $H_0$ in \eqref{eq:H_0} when
+acting on permutationally symmetric (PS) states $\ket\psi\in\M_0$.  To
+perform this decomposition, in this section we simplify products of
+the form $H_0\O\p{\m w,X}\ket\psi$ with $\ket\psi\in\M_0$.  For
+reference, the $M$-body operator $\O\p{\m w,X}$ is defined by
+\begin{align}
+  \O\p{\m w,X} \equiv \sum_{k\in\D_N\p{M}} w_k X_k,
+\end{align}
+where $\m w$ is a dimension-$M$ (i.e.~$M$-index) tensor of scalar
+coefficients $w_k\equiv w_{k_1k_2\cdots k_M}$; $X$ is an $M$-spin
+operator; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is a list of the individual
+spins $k_i\in\ZZ_N$ that the operator $X_k\equiv X_{k_1k_2\cdots k_M}$
+acts on; and
+\begin{align}
+  \D_N\p{M} \equiv
+  \set{ k \in \ZZ_N^M: \t{all entries $k_i$ of $k$ are distinct } },
+\end{align}
+is the strictly ``off-diagonal'' part of $\ZZ_N^M$.
 
@@
 
-In order to build familiarity with the problem at hand, we first
-consider the simple case of a single-body perturbation and simplify
-\begin{align}
-  H_0 X_{\m w} \ket\psi
-  = \sum_{\substack{k\in\ZZ_N\\\set{p,q}\in\C_N\p{2}}}
-  h_{pq} w_k \Pi_{pq} X_k \ket\psi.
+To build familiarity with the problem at hand, we first consider the
+simple case of a single-body ($M=1$) case and simplify
+\begin{align}
+  H_0 \O\p{\m w,X} \ket\psi
+  = \f12 \sum_{i\ne j} \sum_k h_{ij} w_k \Pi_{ij} X_k \ket\psi.
   \label{eq:single_body_diagnosis_start}
@@
 The sum in \eqref{eq:single_body_diagnosis_start} has terms with
-$k\notin\set{p,q}$, and terms with $k\in\set{p,q}$.  In the case of
-$k\notin\set{p,q}$, the permutation operator $\Pi_{pq}$ commutes with
-$X_k$ and annihilates on $\ket\psi$, allowing us to simplify
-\begin{align}
-  \sum_{k\in\ZZ_N} \sum_{\substack{\set{p,q}\in\C_N\p{2}\\k\notin\set{p,q}}}
-  h_{pq} w_k \Pi_{pq} X_k \ket\psi
-  = E_0 X_{\m w} \ket\psi
-  - \sum_{k\in\ZZ_N} \sum_{\substack{\set{p,q}\in\C_N\p{2}\\k\in\set{p,q}}}
-  h_{pq} w_k X_k \ket\psi,
-\end{align}
-where $E_0 = \sum_{\set{p,q}\in\C_N\p{2}} h_{pq}$ is the interaction
-energy of PS states $\ket\psi\in\M_0$, and
-\begin{align}
-  \sum_{\substack{\set{p,q}\in\C_N\p{2}\\k\in\set{p,q}}} h_{pq}
-  = \f12 \sum_{\substack{p,q\in\ZZ_N\\k\in\set{p,q}}} h_{pq}
-  = \f12 \sum_{p\in\ZZ_N} h_{pk} + \f12 \sum_{q\in\ZZ_N} h_{kq}
+$k\in\set{i,j}$, and terms with $k\notin\set{i,j}$.  In the case of
+$k\notin\set{i,j}$, the permutation operator $\Pi_{ij}$ commutes with
+$X_k$ and annihilates on $\ket\psi$, and we can replace the sum
+\begin{align}
+  \sum_{k\notin\set{i,j}} \to \sum_k - \sum_{k\in\set{i,j}},
+\end{align}
+allowing us to simplify
+\begin{align}
+  \f12 \sum_{i\ne j} \sum_{k\notin\set{i,j}}
+  h_{ij} w_k \Pi_{ij} X_k \ket\psi
+  = E_0 \O\p{\m w,X} \ket\psi
+  - \f12 \sum_{i\ne j} \sum_{k\in\set{i,j}} h_{ij} w_k X_k \ket\psi,
+\end{align}
+where $E_0=\frac12\sum_{i\ne j}h_{ij}$ is the interaction energy of PS
+states $\ket\psi\in\M_0$.  Making the replacement
+\begin{align}
+  \sum_{i\ne j} \sum_{k\in\set{i,j}}
+  \to \sum_k \sum_{\substack{i\ne j\\\set{i,j}\ni k}},
+\end{align}
+we can simplify
+\begin{align}
+  \f12 \sum_{\substack{i\ne j\\\set{i,j}\ni k}} h_{ij}
+  = \f12 \sum_i h_{ik} + \f12 \sum_j h_{kj}
   = h_k,
   &&
-  h_k \equiv \sum_{p\in\ZZ_N} h_{pk},
-\end{align}
-which implies
-\begin{align}
-  \sum_{k\in\ZZ_N} \sum_{\substack{\set{p,q}\in\C_N\p{2}\\k\notin\set{p,q}}}
-  h_{pq} w_k X_k \ket\psi
-  = E_0 X_{\m w} \ket\psi
-  - \sum_{k\in\ZZ_N} h_k w_k X_k \ket\psi.
+  h_k \equiv \sum_i h_{ik},
+\end{align}
+which implies that the terms in \eqref{eq:single_body_diagnosis_start}
+with $k\notin\set{i,j}$ are
+\begin{align}
+  \f12 \sum_{i\ne j} \sum_{k\notin\set{i,j}}
+  h_{ij} w_k \Pi_{ij} X_k \ket\psi
+  = E_0 \O\p{\m w,X} \ket\psi - \sum_k h_k w_k X_k \ket\psi.
 \end{align}
 The terms in \eqref{eq:single_body_diagnosis_start} with
-$k\in\set{p,q}$, meanwhile, are
-\begin{align}
-  \sum_{\substack{\set{p,q}\in\C_N\p{2}\\k\in\set{p,q}}}
-  h_{pq} w_k \Pi_{pq} X_k \ket\psi
-  = \sum_{k,q\in\ZZ_N} h_{kq} w_k X_q \ket\psi,
+$k\in\set{i,j}$, meanwhile, are
+\begin{align}
+  \f12 \sum_{\substack{i\ne j\\k\in\set{i,j}}}
+  h_{ij} w_k \Pi_{ij} X_k \ket\psi
+  = \sum_{j,k} h_{kj} w_k X_j \ket\psi.
 \end{align}
@@
 \begin{align}
-  H_0 X_{\m w} \ket\psi
-  = E_0 X_{\m w} \ket\psi
-  + \sum_{p\in\ZZ_N} \p{\sum_{q\in\ZZ_N} h_{pq} w_q - h_p w_p}
-  X_p \ket\psi.
+  H_0 \O\p{\m w,X} \ket\psi
+  = E_0 \O\p{\m w,X} \ket\psi
+  + \sum_k \sp{\sum_j h_{kj} w_j - h_k w_k} X_k \ket\psi.
   \label{eq:single_body_diagnosis_end}
 \end{align}
-The action of the single-body perturbation $X_{\m w}$ on a
+The action of the single-body perturbation $\O\p{\m w,X}$ on a
 permutationally symmetric state therefore generates an eigenstate of
-$H_0$ with energy $E_0+\Delta$ if the vector
-$\v{\m w}\equiv\sum_{k\in\ZZ_N} w_k \ket{k}$ satisfies the eigenvalue
-equation
-\begin{align}
-  \p{\m h - \diag\v h} \c \v{\m w} = \Delta \v{\m w},
+$H_0$ with energy $E_0+\Delta$ if the vector $\m w=\sum_k w_k \ket{k}$
+satisfies the eigenvalue equation
+\begin{align}
+  \p{\m h - \diag\v h} \c \m w = \Delta \m w,
   \label{eq:single_body_eig}
 \end{align}
-where $\m h\equiv\sum_{p,q\in\ZZ_N} h_{pq} \op{p}{q}$ is a matrix of
-all couplings $h_{pq}$; the vector
-$\v h \equiv \sum_{p\in\ZZ_N} h_p \ket{p} = \sum_{p,q\in\ZZ_N} h_{pq}
-\ket{p}$ is the sum of all columns of $\m h$; and $\diag\v h$ is a
-matrix with $\v h$ on the diagonal and zeroes everywhere else.
-
-When the interaction Hamiltonian $H_0$ is translationally invariant,
-the single-body eigenvalue problem in \eqref{eq:single_body_eig} is
-solvable analytically.  In this case, the couplings $h_{pq}$ depend
-only on the separation $\abs{p-q}$, so eigenvectors of $\m h$ are
-therefore plane waves of the form
-\begin{align}
-  \v{\m w}_k \equiv \sum_{p\in\ZZ_L^D} e^{ip\c k} \ket{p},
+where $\m h\equiv\sum_{i,j} h_{ij} \op{i}{j}$ is a matrix of all
+couplings $h_{ij}$; the vector
+$\v h \equiv \sum_{i,j} h_{ij} \ket{i} = \sum_i h_i \ket{i}$ is the
+sum of all columns of $\m h$; and $\diag\v h$ is a matrix with $\v h$
+on the diagonal and zeroes everywhere else.  When the interaction
+Hamiltonian $H_0$ is translationally invariant, the single-body
+eigenvalue problem in \eqref{eq:single_body_eig} is solvable
+analytically.  In this case, the couplings $h_{ij}$ depend only on the
+separation $\abs{i-j}$, so eigenvectors of $\m h$ are therefore plane
+waves of the form
+\begin{align}
+  \m w_k \equiv \sum_{d\in\ZZ_L^D} e^{\i d\c k} \ket{d},
 \end{align}
 where on a $D$-dimensional periodic lattice of $N=L^D$ spins, lattice
-sites are indexed by vectors $p\in\ZZ_L^D$, and wavenumbers take on
+sites are indexed by vectors $d\in\ZZ_L^D$, and wavenumbers take on
 values $k\in\ZZ_L^D\times2\pi/L$.  The corresponding eigenvalues of
@@
 \begin{align}
-  \m h\c{\m w}_k
-  = \sum_{p,q\in\ZZ_L^D} h_{pq} e^{iq\c k} \ket{p}
-  = \sum_{p,d\in\ZZ_L^D} h_{p,p+d} e^{i\p{p+d}\c k} \ket{p}
-  = \sum_{d\in\ZZ_L^D} h_{0,d} \cos\p{d\c k} \v{\m w}_k,
-\end{align}
-where the imaginary contributions vanish because $h_{0,d}=h_{0,-d}$.
-The remainder of \eqref{eq:single_body_eig} that we need to sort out
-is $\diag\v h$, where all
-$h_p=\sum_{q\in\ZZ_L^D}h_{pq}=\sum_{d\in\ZZ_L^D}h_{0,d}$ are equal,
-which implies that $\diag\v h=\sum_{d\in\ZZ_L^D}h_{0,d}$ is a scalar.
-We thus find that
-\begin{align}
-  \p{\m h - \diag\v h}\c\v{\m w}_k = \Delta_k \v{\m w}_k,
+  \m h \c \m w_k
+  = \sum_{c,d\in\ZZ_L^D} h_{cd} e^{\i d\c k} \ket{c}
+  = \sum_{c,d\in\ZZ_L^D} h_{c,c+d} e^{\i\p{c+d}\c k} \ket{c}
+  = \sum_{d\in\ZZ_L^D} h_{0,d} \cos\p{d\c k} \m w_k,
+\end{align}
+where the imaginary contributions vanish in the sum over $d$ because
+$h_{0,d}=h_{0,-d}$.  The remainder of \eqref{eq:single_body_eig} that
+we need to sort out is $\diag\v h$, where all
+$h_i = \sum_{i,j}h_{ij} = \sum_d h_{0,d}$ are equal, which implies
+that $\diag\v h = \sum_d h_{0,d}$ is a scalar.  We thus find that
+\begin{align}
+  \p{\m h - \diag\v h} \c \m w_k = \Delta_k \m w_k,
   &&
@@
 \begin{align}
-  H_0 \tilde X_{\m w} \ket\psi
-  = \sum_{\substack{k\in\D_N\p{M}\\\set{p,q}\in\C_N\p{2}}}
-  h_{pq} w_k \Pi_{pq} X_k \ket\psi,
+  H_0 \O\p{\m w,X} \ket\psi
+  = \sum_{\substack{k\in\D_N\p{M}\\\set{i,j}\in\C_N\p{2}}}
+  h_{ij} w_k \Pi_{ij} X_k \ket\psi,
   \label{eq:diagnosis_start}
 \end{align}
-where to avoid factors of $M!$ in this section, we define
-$\tilde X_{\m w}\equiv M! X_{\m w}$.  The sum in
-\eqref{eq:diagnosis_start} has terms with $p,q\notin k$, terms with
-$p,q\in k$, and terms with only one of $p$ or $q\in k$.  The
-permutation operator $\Pi_{pq}$ acts trivially on $X_k\ket\psi$ when
-$p,q\notin k$, allowing us to simplify
-\begin{align}
-  \sum_{\substack{k\in\D_N\p{M}\\\set{p,q}\in\C_N\p{2}\\p,q\notin k}}
-  h_{pq} w_k \Pi_{pq} X_k \ket\psi
-  = E_0 \tilde X_{\m w} \ket\psi
+where
+\begin{align}
+  \C_N\p{q} \equiv \set{ s \subset \ZZ_N : \abs{s} = q }
+  \label{eq:choices}
+\end{align}
+is the set of all subsets (``choices'') of $q$ elements from $\ZZ_N$.
+The sum in \eqref{eq:diagnosis_start} has terms with $i,j\notin k$,
+terms with $i,j\in k$, and terms with only one of $i$ or $j\in k$.
+The permutation operator $\Pi_{ij}$ acts trivially on $X_k\ket\psi$
+when $i,j\notin k$, and we can replace the sum
+\begin{align}
+  \sum_{\substack{\set{i,j}\in\C_N\p{2}\\i,j\notin k}}
+  \to \sum_{\set{i,j}\in\C_N\p{2}}
+  - \sum_{\substack{\set{i,j}\in\C_N\p{2}\\i\in k~\t{or}~j\in k}},
+\end{align}
+allowing us to simplify
+\begin{align}
+  \sum_{k\in\D_N\p{M}} \sum_{\substack{\set{i,j}\in\C_N\p{2}\\i,j\notin k}}
+  h_{ij} w_k \Pi_{ij} X_k \ket\psi
+  = E_0 \O\p{\m w,X} \ket\psi
   - \sum_{k\in\D_N\p{M}} h^{\t{D}}_k w_k X_k \ket\psi,
 \end{align}
-where $E_0 = \sum_{\set{p,q}\in\C_N\p{2}} h_{pq}$ is the interaction
+where $E_0 = \sum_{\set{i,j}\in\C_N\p{2}} h_{ij}$ is the interaction
 energy of PS states $\ket\psi\in\M_0$, and
@@
   h^{\t{D}}_k \equiv
-  \sum_{\substack{\set{p,q}\in\C_N\p{2}\\p\in k~\t{or}~q\in k}} h_{pq}
-  = \sum_{p\in k} h_p - \sum_{\set{a,b}\in\C_M\p{2}} h_{k_ak_b},
-  &&
-  h_p \equiv \sum_{q\in\ZZ_N} h_{pq}.
+  \sum_{\substack{\set{i,j}\in\C_N\p{2}\\i\in k~\t{or}~j\in k}} h_{ij}
+  = \sum_{i\in k} h_i - \sum_{\set{a,b}\in\C_M\p{2}} h_{k_ak_b},
+  &&
+  h_i \equiv \sum_j h_{ij},
   \label{eq:multi_body_op_diag_element}
 \end{align}
-Defining a linear operator $\m h_{\t{D}}$ on the space of
-dimension-$M$ tensors, whose action on $\m w$ is a new dimension-$M$
-tensor with entries
-\begin{align}
-  \m h_{\t{D}}\p{\m w}_k \equiv h^{\t{D}}_k w_k,
+with $k_a,k_b\in k\equiv\p{k_1,k_2,\cdots,k_M}$.  Defining a linear
+operator $\hat{\m h}_{\t{D}}$ on the space of dimension-$M$ tensors,
+whose action on $\m w$ is a new dimension-$M$ tensor with entries
+\begin{align}
+  \p{\hat{\m h}_{\t{D}}\c\m w}_k \equiv h^{\t{D}}_k w_k,
   \label{eq:multi_body_op_diag}
@@
 we can write the terms in \eqref{eq:diagnosis_start} with
-$p,q\notin k$ as
-\begin{align}
-  \sum_{\substack{k\in\D_N\p{M}\\\set{p,q}\in\C_N\p{2}\\p,q\notin k}}
-  h_{pq} w_k \Pi_{pq} X_k \ket\psi
-  = E_0 \tilde X_{\m w} \ket\psi
-  - \sum_{k\in\D_N\p{M}} \m h_{\t{D}}\p{\m w}_k X_k \ket\psi.
-\end{align}
-The terms in \eqref{eq:diagnosis_start} with $p,q\in k$ are
-\begin{align}
-  \sum_{\substack{k\in\D_N\p{M}\\\set{p,q}\in\C_N\p{2}\\p,q\in k}}
-  h_{pq} w_k \Pi_{pq} X_k \ket\psi
+$i,j\notin k$ as
+\begin{align}
+  \sum_{k\in\D_N\p{M}} \sum_{\substack{\set{i,j}\in\C_N\p{2}\\i,j\notin k}}
+  h_{ij} w_k \Pi_{ij} X_k \ket\psi
+  = E_0 \O\p{\m w,X} \ket\psi
+  - \sum_{k\in\D_N\p{M}} \p{\hat{\m h}_{\t{D}}\c\m w}_k X_k \ket\psi.
+\end{align}
+The terms in \eqref{eq:diagnosis_start} with $i,j\in k$ are
+\begin{align}
+  \sum_{k\in\D_N\p{M}} \sum_{\substack{\set{i,j}\in\C_N\p{2}\\i,j\in k}}
+  h_{ij} w_k \Pi_{ij} X_k \ket\psi
   = \sum_{k\in\D_N\p{M}} \sum_{\substack{\set{a,b}\in\C_M\p{2}}}
@@
 $w_{k_{a\lra b}} X_k$.  To write down these terms succinctly, we
-define a linear operator $\m h_{\t{P}}$ on the space of dimension-$M$
-tensors, whose action on $\m w$ is a new dimension-$M$ tensor with
-entries
-\begin{align}
-  \m h_{\t{P}}\p{\m w}_k
+define a linear operator $\hat{\m h}_{\t{P}}$ on the space of
+dimension-$M$ tensors, whose action on $\m w$ is a new dimension-$M$
+tensor with entries
+\begin{align}
+  \p{\hat{\m h}_{\t{P}} \c \m w}_k
   \equiv \sum_{\substack{\set{a,b}\in\C_M\p{2}}}
@@
 \begin{align}
-  \sum_{\substack{k\in\D_N\p{M}\\\set{p,q}\in\C_N\p{2}\\p,q\in k}}
-  h_{pq} w_k \Pi_{pq} X_k \ket\psi
-  = \sum_{k\in\D_N\p{M}} \m h_{\t{P}}\p{\m w}_k X_k \ket\psi.
+  \sum_{k\in\D_N\p{M}} \sum_{\substack{\set{i,j}\in\C_N\p{2}\\i,j\in k}}
+  h_{ij} w_k \Pi_{ij} X_k \ket\psi
+  = \sum_{k\in\D_N\p{M}} \p{\hat{\m h}_{\t{P}} \c \m w}_k X_k \ket\psi.
 \end{align}
 Finally, the terms in \eqref{eq:diagnosis_start} with only one of
-$p\in k$ or $q\in k$ are
-\begin{align}
-  \sum_{k\in\D_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}}
-  \sum_{q\in k} h_{pq} w_k  \Pi_{pq} X_k \ket\psi
-  = \sum_{k\in\D_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}}
-  \sum_{a\in\ZZ_M} h_{pk_a} w_{k_{a:p}} X_k \ket\psi,
-\end{align}
-where $k_{a:p}$ denotes a list that is equal to $k$ but with $k_a$
-replaced by $p$, and we have relabeled $k_a$ and $p$ to change
-$w_k X_{k_{a:p}}$ to $w_{k_{a:p}} X_k$.  For each fixed $a$, the sum
-over $p$ above is essentially a contraction of the matrix $\m h$ with
+$i\in k$ or $j\in k$ are
+\begin{align}
+  \sum_{k\in\D_N\p{M}} \sum_{i\notin k} \sum_{j\in k}
+  h_{ij} w_k \Pi_{ij} X_k \ket\psi
+  = \sum_{k\in\D_N\p{M}} \sum_{i\notin k}
+  \sum_{a\in\ZZ_M} h_{ik_a} w_{k_{a:i}} X_k \ket\psi,
+\end{align}
+where $k_{a:i}$ denotes a list that is equal to $k$ but with $k_a$
+replaced by $i$, and we have relabeled $k_a$ and $i$ to change
+$w_k X_{k_{a:i}}$ to $w_{k_{a:i}} X_k$.  For each fixed $a$, the sum
+over $i$ above is essentially a contraction of the matrix $\m h$ with
 the $a$-th index of the tensor $\m w$.  Defining a linear operator
-$\m h_a$ on the space of dimension-$M$ tensors, whose action on $\m w$
-is a new dimension-$M$ tensor with entries
-\begin{align}
-  \m h_a\p{\m w}_k
-  \equiv \sum_{\substack{p\in\ZZ_N\\p\notin k}} h_{pk_a} w_{k_{a:p}}
-  = \sum_{\substack{p\in\ZZ_N\\p\notin k}}
-  h_{pk_a} w_{k_1 k_2 \cdots k_{a-1} p k_{a+1} \cdots k_M},
+$\hat{\m h}_a$ on the space of dimension-$M$ tensors, whose action on
+$\m w$ is a new dimension-$M$ tensor with entries
+\begin{align}
+  \p{\hat{\m h}_a\c\m w}_k
+  \equiv \sum_{i\notin k} h_{ik_a} w_{k_{a:i}}
+  = \sum_{i\notin k}
+  h_{ik_a} w_{k_1 k_2 \cdots k_{a-1} i k_{a+1} \cdots k_M},
   \label{eq:multi_body_op_contract}
@@
 \begin{align}
-  \sum_{k\in\D_N\p{M}} \sum_{\substack{p\in\ZZ_N\\p\notin k}}
-  \sum_{q\in k} h_{pq} w_k  \Pi_{pq} X_k \ket\psi
+  \sum_{k\in\D_N\p{M}} \sum_{i\notin k} \sum_{j\in k}
+  h_{ij} w_k \Pi_{ij} X_k \ket\psi
   = \sum_{k\in\D_N\p{M}} \sum_{a\in\ZZ_M}
-  \m h_a\p{\m w}_k X_k \ket\psi.
-\end{align}
-Altogether, we find that
-\begin{align}
-  H_0 X_{\m w} \ket\psi
-  = E_0 X_{\m w} \ket\psi + X_{\hat{\m h}\p{\m w}} \ket\psi,
-\end{align}
-where
-\begin{align}
-  \hat{\m h}\p{\m w}
-  \equiv \sum_{a\in\ZZ_M} \m h_a\p{\m w}
-  + \m h_{\t{P}}\p{\m w} - \m h_{\t{D}}\p{\m w},
-\end{align}
-is defined in terms of the tensors $\m h_a\p{\m w}$,
-$\m h_{\t{P}}\p{\m w}$, and $\m h_{\t{D}}\p{\m w}$, whose entries are
-respectively provided in \eqref{eq:multi_body_op_contract},
-\eqref{eq:multi_body_op_perm}, and \eqref{eq:multi_body_op_diag}.  The
-action of the $M$-body perturbation $X_{\m w}$ on a permutationally
-symmetric state therefore generates an eigenstate of $H_0$ with energy
-$E_0+\Delta$ if the coefficient tensor $\m w$ satisfies the
-generalized eigenvalue equation $\hat{\m h}\p{\m w} = \Delta \m w$,
-which can be written in the form
-\begin{align}
-  \hat{\m h} \c \v{\m w} = \Delta \v{\m w},
-  &&
-  \v{\m w} \equiv \sum_{k\in\D_N\p{M}} w_k \ket{k},
+  \p{\hat{\m h}_a\c\m w}_k X_k \ket\psi.
+\end{align}
+Altogether, defining the combined operator
+\begin{align}
+  \hat{\m h}
+  \equiv \sum_{a\in\ZZ_M} \m h_a + \m h_{\t{P}} - \m h_{\t{D}},
+\end{align}
+we find that
+\begin{align}
+  H_0 \O\p{\m w,X} \ket\psi
+  = E_0 \O\p{\m w,X} \ket\psi +
+  \O\p{\hat{\m h}\c\m w, X} \ket\psi
+  = \sum_\Delta \p{E_0+\Delta} \O\p{\m w_\Delta, X} \ket\psi,
+\end{align}
+where $\m w_\Delta$ is the projection of $\m w$ onto the eigenspace of
+$\hat{\m h}$ with eigenvalue $\Delta$, i.e.
+\begin{align}
+  \hat{\m h} \c \m w_\Delta = \Delta \m w_\Delta,
   \label{eq:multi_body_problem}
 \end{align}
-where
+with
 \begin{align}
   \hat{\m h} \equiv \sum_{k\in\D_N\p{M}} \ket{k}
-  \sp{\sum_{a\in\ZZ_M} \sum_{\substack{p\in\ZZ_N\\p\notin k}}
-    h_{pk_a} \bra{k_{a:p}}
+  \sp{\sum_{a\in\ZZ_M} \sum_{\substack{i\in\ZZ_N\\i\notin k}}
+    h_{ik_a} \bra{k_{a:i}}
     + \sum_{\set{a,b}\in\C_M\p{2}} h_{k_ak_b} \bra{k_{a\lra b}}
@@
 \end{align}
+For reference, here $k_{a:i}$ is equal to
+$k\equiv\p{k_1,k_2,\cdots,k_M}$ but with $k_a$ replaced by $i$;
+$\C_M\p{2}$ is defined in \eqref{eq:choices}; $k_{a\lra b}$ is equal
+to $k$ but with $k_a$ and $k_b$ swapped; and $h^{\t{D}}_k$ is defined
+in \eqref{eq:multi_body_op_diag_element}.
+
 If the coefficient tensor $\m w$ is permutationally symmetric, meaning
@@
 over $k\in\C_N\p{M}$, and replace vectors
-$\ket{\p{k_1,k_2,\cdots,k_M}}\to\ket{\set{k_1,k_2,\cdots,k_M}}$, such
-that e.g.~$\ket{k_{a\lra b}}=\ket{k}$.  These replacements reduce the
-size of $\hat{\m h}$ from $\abs{\D_N\p{M}}\times\abs{\D_N\p{M}}$ to
+$\ket{k_1,k_2,\cdots,k_M}\to\ket{\set{k_1,k_2,\cdots,k_M}}$, such that
+e.g.~$\ket{k_{a\lra b}}=\ket{k}$.  These replacements reduce the size
+of $\hat{\m h}$ from $\abs{\D_N\p{M}}\times\abs{\D_N\p{M}}$ to
 $\abs{\C_N\p{M}}\times\abs{\C_N\p{M}}$, where
-$\abs{\D_N\p{M}}=M!\times{N\choose M}=\prod_{j=0}^{M-1}\p{N-j}$ and
+$\abs{\D_N\p{M}}=\prod_{j=0}^{M-1}\p{N-j}=M!\times{N\choose M}$ and
 $\abs{\C_N\p{M}}={N\choose M}$.  We discuss further reduction of the
@@
   \hat{\m h} \stackrel{M=1}{=} \sum_{k\in\ZZ_N} \ket{k}
-  \sp{\sum_{\substack{p\in\ZZ_N\\p\ne k}} h_{pk} \bra{p} - h_k \bra{k}}
+  \sp{\sum_{i\in\ZZ_N} h_{ik} \bra{i} - h_k \bra{k}}
   = \m h - \diag\v h,
@@
   \hat{\m h} \stackrel{M=2}{=}
-  \sum_{\p{k,\ell}\in\D_N\p{2}} \ket{k,\ell}
-  \sp{\sum_{\substack{p\in\ZZ_N\\p\notin\set{k,\ell}}}
-    \p{ h_{pk} \bra{p,\ell} + h_{p\ell} \bra{k,p} }
-    + h_{k\ell} \bra{\ell,k}
-    - \p{h_k + h_\ell - h_{k\ell}} \bra{k,\ell}}.
+  \sum_{\p{k,\ell}\in\D_N\p{2}} \ket{k\ell}
+  \sp{\sum_{\substack{i\in\ZZ_N\\i\notin\set{k,\ell}}}
+    \p{ h_{ik} \bra{i\ell} + h_{i\ell} \bra{ki} }
+    + h_{k\ell} \bra{\ell k}
+    - \p{h_k + h_\ell - h_{k\ell}} \bra{k\ell}}.
 \end{align}
@@
 Solving the eigenvalue problem in \eqref{eq:multi_body_problem} to
-find $M$-body operators $X_{\m w}$ that generate states of definite
-interaction energy nominally requires diagonalizing the matrix
-$\hat{\m h}$ with dimensions $N!_M\times N!_M\sim N^M\times N^M$,
-where $N!_M=\prod_{j=0}^{M-1}\p{N-j}$ denotes a falling factorial.
-Spatial symmetries of the interaction Hamiltonian $H_0$, however,
-generally endow $\hat{\m h}$ with a block-diagonal structure that can
-be used to reduce the complexity of this eigenvalue problem.
-Translational invariance, for example, can reduce this problem to that
-of diagonalizing a matrix with dimensions
-$\sim N^{M-1}\times N^{M-1}$, and isotropy can further reduce the
-problem size by factors that depend on the lattice geometry.  The
-complexity of diagonalizing a matrix with dimensions $d\times d$ is
-$O\p{d^3}$, so even a reduction of $d$ by a factor of 2 yields
-practically significant gains (a factor of $2^3=8$) in the runtime of
-diagonalization.  Here, we provide a general prescription for
-exploiting the spatial symmetries of $H_0$ to simplify the eigenvalue
-problem in \eqref{eq:multi_body_problem}.
+find $M$-body operators $\O\p{\m w_\Delta,X}$ that generate states of
+definite interaction energy nominally requires diagonalizing the
+matrix $\hat{\m h}$ with dimensions $\sim N^M\times N^M$.  Spatial
+symmetries of the interaction Hamiltonian $H_0$, however, generally
+endow $\hat{\m h}$ with a block-diagonal structure that can be used to
+reduce the complexity of this eigenvalue problem.  Translational
+invariance, for example, can reduce this problem to that of
+diagonalizing a matrix with dimensions $\sim N^{M-1}\times N^{M-1}$,
+and isotropy can further reduce the problem size by factors that
+depend on the lattice geometry.  The complexity of diagonalizing a
+matrix with dimensions $d\times d$ is $O\p{d^3}$, so even a reduction
+of $d$ by a factor of 2 yields practically significant gains (a factor
+of $2^3=8$) in the runtime of diagonalization.  Here, we provide a
+general prescription for exploiting the spatial symmetries of $H_0$ to
+simplify the eigenvalue problem in \eqref{eq:multi_body_problem}.
 
@@
 isotropic system, for example, each equivalence class $s\in\S_2$ would
-consist of all pairs $\p{p,q}\in\D_N\p{2}$ that are a fixed distance
-$\abs{p-q}=d_s$ apart,
-i.e.~$s=\set{\p{p,q}\in\D_N\p{2}:\abs{p-q}=d_s}$.  If the tensor
+consist of all pairs $\p{i,j}\in\D_N\p{2}$ that are a fixed distance
+$\abs{i-j}=d_s$ apart,
+i.e.~$s=\set{\p{i,j}\in\D_N\p{2}:\abs{i-j}=d_s}$.  If the tensor
 $\m w$ is invariant under arbitrary permutations of its tensor
@@
 
-For simplicity, we first assume that the $M$-body operator $X_{\m w}$
-obeys the same spatial symmetries as $H_0$.  For any equivalence class
-$s\in\S_M$, we then define a uniform superposition $\ket{s}$ over all
-$k\in s$, as well as a projector $\P_\S$ onto the space of uniform
-superpositions:
+For simplicity, we first assume that the $M$-body operator
+$\O\p{\m w,X}$ obeys the same spatial symmetries as $H_0$.  For any
+equivalence class $s\in\S_M$, we then define a uniform superposition
+$\ket{s}$ over all $k\in s$, as well as a projector $\P_\S$ onto the
+space of uniform superpositions:
 \begin{align}
@@
 where $\abs{s}$ is the number of elements in $s$.  In turn, we expand
-the coefficient vector $\v{\m w}$ in this symmetrized basis:
-\begin{align}
-  \v{\m w} = \P_{\S_M} \v{\m w}
+a vector of $M$-body operator coefficients
+$\m w=\sum_{k\in\D_N\p{M}}w_k\ket{k}$ in this symmetrized basis:
+\begin{align}
+  \m w = \P_{\S_M} \m w
   = \sum_{s\in\S_M} \sum_{k\in s} w_k \ket{s} \bk{s|k}
@@
 \begin{align}
-  \P_{\S_M} \hat{\m h} \P_{\S_M} \c \v{\m w} = \Delta \v{\m w}.
-\end{align}
-More generally, even if the couplings $\v{\m w}$ of $X_{\m w}$ do not
+  \P_{\S_M} \hat{\m h} \P_{\S_M} \c \m w = \Delta \m w.
+\end{align}
+More generally, even if the couplings $\m w$ of $\O\p{\m w,X}$ do not
 obey the spatial symmetries as $H_0$, these symmetries generally endow
@@
 \begin{align}
-  \m X_{\m w}
-  \equiv \prod_{\alpha\in\ZZ_p} \f1{M_\alpha!}
-  \sum_{k_\alpha\in\D_N\p{M_\alpha}}
-  w_\alpha\p{k_\alpha} X_\alpha\p{k_\alpha}
-  = \f1{\m M!} \sum_{k\in\D_N\p{\m M}} \prod_{\alpha\in\ZZ_p}
+  \Q\p{\m w,\m X}
+  \equiv \prod_{\alpha\in\ZZ_p} \O\p{w_\alpha,X_\alpha}
+  = \prod_{\alpha\in\ZZ_p} \sum_{k_\alpha\in\D_N\p{M_\alpha}}
   w_\alpha\p{k_\alpha} X_\alpha\p{k_\alpha},
@@
 \end{align}
-where $\m X\equiv\p{X_1,X_2,\cdots,X_p}$ is a vector of
-$M_\alpha$-spin operators $X_\alpha$;
-$\m w\equiv\p{w_1,w_2,\cdots,w_p}$ is a vector of dimension-$M_\alpha$
-tensors $w_\alpha$; $\m M\equiv\p{M_1,M_2,\cdots,M_p}$ is a vector of
-the dimensions $M_\alpha$;
+where $\m w\equiv\p{w_1,w_2,\cdots,w_p}$ is a vector of
+dimension-$M_\alpha$ tensors $w_\alpha$;
+$\m X\equiv\p{X_1,X_2,\cdots,X_p}$ is a vector of $M_\alpha$-spin
+operators $X_\alpha$;
 $k_\alpha=\p{k_{\alpha,1},k_{\alpha,2},\cdots,k_{\alpha,M_\alpha}}$ is
 a choice of $M_\alpha$ spins that $X_\alpha\p{k_\alpha}$ acts on;
-$w_\alpha\p{k_\alpha}$ is a scalar entry of $w_\alpha$;
+$w_\alpha\p{k_\alpha}$ is a scalar entry of $w_\alpha$; and
 $\D_N\p{M_\alpha}$ is the strictly off-diagonal part of $\ZZ_N^M$,
-defined in \eqref{eq:off_diags};
+defined in \eqref{eq:off_diags}.  For later convenience, we extend the
+definitions of $w_\alpha\p{k_\alpha}=X_\alpha\p{k_\alpha}$ to all
+$k_\alpha\in\ZZ_N^M$ by defining
+$w_\alpha\p{k_\alpha}=X_\alpha\p{k_\alpha}=0$ for any $k_\alpha$ with
+repeated entries.
+
+To simplify $\Q\p{\m w,\m X}$, we first collect terms to express
+\eqref{eq:sym_prod_start} as a sum of products:
+\begin{align}
+  \Q\p{\m w,\m X} = \sum_{k\in\D_N\p{\m M}} \prod_{\alpha\in\ZZ_p}
+  w_\alpha\p{k_\alpha} X_\alpha\p{k_\alpha},
+  \label{eq:sym_prod_sum}
+\end{align}
+where $\m M\equiv\p{M_1,M_2,\cdots,M_p}$ is a vector of the dimensions
+$M_\alpha$; and
 $\D_N\p{\m M}\equiv\bigotimes_{M_\alpha\in\m M}\D_N\p{M_\alpha}$, such
 that each $k\in\D_N\p{\m M}$ has the decomposition
-$k=\p{k_1,k_2,\cdots,k_p}$ with $k_\alpha\in\D_N\p{M_\alpha}$; and
-$\m M!\equiv\prod_{M_\alpha\in\m M}M_\alpha!$ for shorthand.  For
-later convenience, we define
-$w_\alpha\p{k_\alpha}=X_\alpha\p{k_\alpha}=0$ for any $k_\alpha$ with
-repeated entries.
-
-We now classify terms in \eqref{eq:sym_prod_start} by the numbers
+$k=\p{k_1,k_2,\cdots,k_p}$ with $k_\alpha\in\D_N\p{M_\alpha}$.  We
+then classify the terms in \eqref{eq:sym_prod_sum} by the numbers
 $g_S$ of indices shared by all tensors $w_\alpha$ with
@@
 \begin{align}
-  \m X_{\m w} \EQPS \sum_{g\in\G\p{\m M}} \m X_{\m w}\p{g},
+  \Q\p{\m w,\m X} \EQPS \sum_{g\in\G\p{\m M}} \Q_g\p{\m w,\m X},
   \label{eq:sym_prod_diagrams}
@@
 where $\EQPS$ denotes equality under a restriction to $\M_0$, and
-$\m X_{\m w}\p{g}$ is the sum of all terms in
+$\Q_g\p{\m w,\m X}$ is the sum of all terms in
 \eqref{eq:sym_prod_start} that are associated with the diagram $g$.
 
-To find an explicit form of $\m X_{\m w}\p{g}$, we first consider all
+To find an explicit form of $\Q_g\p{\m w,\m X}$, we first consider all
 distinct ways to assign $\abs{g}$ indices to specific tensor factors
@@
 nominally $M_\alpha!$ such choices for a given $\alpha$, and therefore
-$\prod_{\alpha\in\ZZ_p}M_\alpha!=\m M!$ choices total.  The only catch
-is that these choices treat all dots in $g$ as distinct, whereas the
+$\prod_{\alpha\in\ZZ_p}M_\alpha!$ choices total.  The only catch is
+that these choices treat all dots in $g$ as distinct, whereas the
 $g_S$ dots within region $S$ of $g$ are actually indistinguishable.
@@
 \begin{align}
-  \abs{\L\p{g}} = \m M! \times \S\p{g},
-  &&
-  \S\p{g} \equiv \sp{\prod_{S\in\PP^*\p{\ZZ_p}} g_S!}^{-1}.
-  \label{eq:assignment_num}
+  \abs{\L\p{g}} = \sp{\prod_{\alpha\in\ZZ_p}M_\alpha!}
+  \times \sp{\prod_{S\in\PP^*\p{\ZZ_p}} g_S!}^{-1}.
+  \label{eq:symmetry}
 \end{align}
@@
 \begin{align}
-  \m X_{\m w}\p{g} \equiv \f{\S\p{g}}{\abs{\L\p{g}}}
+  \Q_g\p{\m w,\m X} \equiv
   \sum_{\ell\in\L\p{g}} \sum_{k\in\D_N\p{\abs{g}}}
@@
 $k=\p{k_1,k_2,\cdots,k_{\abs{g}}}$ to $\m w$ and $\m X$ according to
-$\ell$.  The specific choice of spins $k$ in $\m X\p{k,\ell}$ does not
-matter after projection onto the PS manifold $\M_0$, so the dependence
-of $\m X\p{k,\ell}$ on $k$ is superfluous, which allows us to write
-\begin{align}
-  \m X_{\m w}\p{g}
-  \EQPS \f{\S\p{g}}{\abs{\L\p{g}}} \sum_{\ell\in\L\p{g}}
-  \m w\p{\ell} \m X\p{\ell},
-  &&
-  \m w\p{\ell} \equiv \sum_{k\in\D_N\p{\abs{g}}} \m w\p{k,\ell}.
-  \label{eq:diagram_terms}
-\end{align}
-If, furthermore, all tensors in $\m w$ (or all operators in $\m X$)
-are {\it permutationally symmetric}, or invariant under arbitrary
-permutations of their tensor factors, then
-\begin{align}
-  \m X_{\m w}\p{g} \stackrel{\t{sym}}{=}_{\t{PS}}
-  \S\p{g} \m w\p{g} \m X\p{g},
+$\ell$.  Upon projection to the PS manifold $\M_0$, this operator
+becomes
+\begin{align}
+  \Q_g\p{\m w,\m X} \equiv
+  \sum_{\ell\in\L\p{g}} \col{\m w}\p{\ell} \mean{\m X}\p{\ell},
+  \label{eq:sum_assignments}
+\end{align}
+where
+\begin{align}
+  \col{\m w}\p{\ell} \equiv \sum_{k\in\D_N\p{\abs{g}}} \m w\p{k,\ell},
+  &&
+  \mean{\m X}\p{\ell}
+  \equiv \f1{\abs{\D_N\p{\abs{g}}}}
+  \sum_{k\in\D_N\p{\abs{g}}} \m X\p{k,\ell},
+\end{align}
+and permutational symmetry implies that the averaging for
+$\mean{\m X}\p{\ell}$ is not actually necessary:
+\begin{align}
+  \mean{\m X}\p{\ell}
+  \EQPS \m X\p{k,\ell} ~\t{for any}~ k\in\D_N\p{\abs{g}}.
+\end{align}
+If, furthermore, all tensors $w_\alpha$ in $\m w$ (or all operators
+$X_\alpha$ in $\m X$) are permutationally symmetric, meaning that they
+are invariant under arbitrary permutations of their tensor factors,
+then
+\begin{align}
+  \Q_g\p{\m w,\m X}
+  \stackrel{\t{sym}}{=}_{\t{PS}} \col{\m w}\p{g} \mean{\m X}\p{g},
+  \label{eq:sum_diagrams}
 \end{align}
@@
 \begin{align}
-  \m w\p{g} \equiv \f1{\abs{\L\p{g}}}
-  \sum_{\ell\in\L\p{g}} \m w\p{\ell}
-  &&
-  \m X\p{g} \equiv \f1{\abs{\L\p{g}}}
-  \sum_{\ell\in\L\p{g}} \m X\p{\ell}.
-  \label{eq:diagram_factors}
-\end{align}
-In summary, the product of operators $\m X_{\m w}$ in
+  \col{\m w}\p{g} \equiv \sum_{\ell\in\L\p{g}} \col{\m w}\p{\ell},
+  &&
+  \mean{\m X}\p{g} \equiv \f1{\abs{\L\p{g}}}
+  \sum_{\ell\in\L\p{g}} \mean{\m X}\p{\ell}.
+\end{align}
+In summary, the product of operators $\Q\p{\m w,\m X}$ in
 \eqref{eq:sym_prod_start} admits the diagrammatic expansion in
-\eqref{eq:sym_prod_diagrams}, as a sum of $\m X_{\m w}\p{g}$ over all
-diagrams $g\in\G\p{\m M}$.  Each term $\m X_{\m w}\p{g}$ is, up to a
-symmetry factor $\S\p{g}$, an average of $\m w\p{\ell} \m X\p{\ell}$
-over all distinct index assignments $\ell\in\L\p{g}$, as in
-\eqref{eq:diagram_terms}.  The scalar $\m w\p{\ell}$ is obtained by
-assigning indices $k=\p{k_1,k_2,\cdots,k_{\abs{g}}}$ to the tensors
-$\m w$ as prescribed by $\ell$, and summing over all
-$k\in\D_N\p{\abs{g}}$, while the $\abs{g}$-spin operator
-$\m X\p{\ell}$ is obtained by contracting the operators in $\m X$ in
-accordance with $\ell$.  If the tensors $\m w$ or operators $\m X$ are
-all permutationally symmetric, then $\m X_{\m w}\p{g}$ factorizes into
-$\S\p{g} \m w\p{g} \m X\p{g}$, where $\m w\p{g}$ and $\m X\p{g}$ are
-respectively averages of $\m w\p{\ell}$ and $\m X\p{\ell}$ over all
-index assignments $\ell\in\L\p{g}$.
+\eqref{eq:sym_prod_diagrams}, as a sum of $\Q_g\p{\m w,\m X}$ over all
+diagrams $g\in\G\p{\m M}$.  Each term $\Q_g\p{\m w,\m X}$ is a sum of
+$\col{\m w}\p{\ell}\mean{\m X}\p{\ell}$ over all distinct index
+assignments $\ell\in\L\p{g}$.  The scalar $\col{\m w}\p{\ell}$ is
+obtained by assigning indices $k=\p{k_1,k_2,\cdots,k_{\abs{g}}}$ to
+the tensors $\m w$ as prescribed by $\ell$, and summing over all
+values of $k\in\D_N\p{\abs{g}}$.  The collective operator
+$\mean{\m X}\p{\ell}$ is similarly constructed with an average over
+all $k\in\D_N\p{\abs{g}}$, rather than a sum.  If the tensors $\m w$
+or operators $\m X$ are all permutationally symmetric, then
+$\Q_g\p{\m w,\m X}$ factorizes into $\col{\m w}\p{g}\mean{\m X}\p{g}$,
+where $\col{\m w}\p{g}$ and $\mean{\m X}\p{g}$ are respectively the
+sum and average of $\col{\m w}\p{\ell}$ and $\mean{\m X}\p{\ell}$ over
+all index assignments $\ell\in\L\p{g}$.
+
+As a final point, we note that the choice to sum (average) for the
+scalar (operator) content of $\Q_g\p{\m w,\m X}$ in
+\eqref{eq:sum_assignments} and \eqref{eq:sum_diagrams} was arbitrary,
+and we could just as well have chosen to perform an average (sum) for
+the scalar (operator) content.  For later convenience, we therefore
+define the {\it mixed} sums and averages
+\begin{align}
+  \ut{\m w}\p{g} \equiv \f1{\abs{\L\p{g}}}
+  \sum_{\ell\in\L\p{g}} \col{\m w}\p{\ell}
+  = \f1{\abs{\L\p{g}}} \sum_{\ell\in\L\p{g}}
+  \sum_{k\in\D_N\p{\abs{g}}} \m w\p{k,\ell},
+\end{align}
+\begin{align}
+  \ot{\m X}\p{g} \equiv \sum_{\ell\in\L\p{g}} \mean{\m X}\p{\ell}
+  \EQPS \sum_{\ell\in\L\p{g}} \m X\p{k,\ell}
+  ~\t{for any}~ k\in\D_N\p{\abs{g}}.
+\end{align}
+If all tensors $w_\alpha$ in $\m w$ (or all operators $X_\alpha$ in
+$\m X$) are permutationally symmetric, then
+\begin{align}
+  \Q_g\p{\m w,\m X} \stackrel{\t{sym}}{=}_{\t{PS}}
+  \ut{\m w}\p{g} \ot{\m X}\p{g}.
+  \label{eq:sum_diagrams_alt}
+\end{align}
 
@@
 Diagrams such as \eqref{eq:venn_diagram} play a central role in
-simplifying products of multi-body operators, through the expansion in
-\eqref{eq:sym_prod_diagrams}.  The purpose of diagrams is to
-succinctly represent the scalar content of each term in this
-expansion.  Here, we define these diagrams as formal objects that
-enable us to systematically perform otherwise intractable
-calculations.  We then derive rules to ``reduce'' these diagrams in
-such a way as to minimize the computational cost of their numerical
-evaluation.
-
-All diagrams consist of $p$ circles, with each circle $\alpha\in\ZZ_p$
-labeled by some tensor $w_\alpha$, as in \eqref{eq:venn_diagram} (or
-below).  The simplest diagrams have $g_S$ filled dots (``$\bullet$'')
-at the intersection of the circles $\alpha\in S\subset\ZZ_p$, and are
-simply equal to the scalar $\S\p{g} \m w\p{g}$ defined by
-\eqref{eq:assignment_num}, \eqref{eq:diagram_terms}, and
-\eqref{eq:diagram_factors}.  If all tensors $w_\alpha$ are
-permutationally symmetric, for example, then
-\begin{align}
-  \diagram{example_123}
-  = \f1{2!} \sum_{\p{a,b,c,d,e}\in\D_N\p{5}}
+simplifying products of multi-body operators, $\Q\p{\m w,\m X}$
+through the expansion of these products in
+\eqref{eq:sym_prod_diagrams} as sum over $\Q_g\p{\m w,\m X}$.  Here,
+we formalize a diagrammatic representation of the scalars
+$\col{\m w}\p{\ell}$ and $\ut{\m w}\p{g}$ that appear in
+$\Q_g\p{\m w,\m X}$, which enables us to perform otherwise intractable
+calculations.  We then derive rules to ``reduce'' these scalars so as
+to minimize the computational cost of their numerical evaluation.
+
+A diagram $g\in\G\p{\m M}$ with $\m M=\p{M_1,M_2,\cdots,M_p}$ consists
+of $p$ circles, and $g_S$ dots at the intersection of circles
+$\alpha\in S\subset\ZZ_p$.  By slight abuse of notation, labeling each
+circle $\alpha$ by a tensor $w_\alpha$ defines the scalar
+$g\p{\m w}\equiv\ut{\m w}\p{g}$.  If, for example, all tensors
+$w_\alpha$ are all permutationally symmetric, then
+\begin{align}
+  \diagram{example_w123}
+  = \sum_{\p{a,b,c,d,e}\in\D_N\p{5}}
   w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}.
@@
 When the tensors $\m w$ are not all permutationally symmetric, one can
-additionally equip a diagram with a specific index assignment
+additionally equip a diagram $g$ with a specific index assignment
 $\ell\in\L\p{g}$, in which case the diagram is equal to
-$\S\p{g} \m w\p{\ell}$.  Diagrams that are not labeled with a specific
-index assignment $\ell$ are then an average over all $\ell\in\L\p{g}$.
-To keep mathematical expressions simple, we will use symmetric tensors
-in all examples in this section, but note that our discussions apply
-just as well to the asymmetric case.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Reducing diagrams}
+$\ell\p{\m w}\equiv\col{\m w}\p{\ell}$.  Diagrams that are not
+equipped with a specific index assignment $\ell$ are then an average
+over all $\ell\in\L\p{g}$,
+i.e.~$g\p{\m w}\equiv\sum_{\ell\in\L\p{g}} \ell\p{\m w} /
+\abs{\L\p{g}}$.  To keep mathematical expressions simple, we will use
+symmetric tensors in all examples in this section, but our discussions
+apply just as well to the asymmetric case.
 
@@
   \diagram{example_o}
-  \propto \sum_{\p{a,b,e}\in\D_N\p{3}} \sum_{c,d,e\in\ZZ_N}
-  v\p{a,b,c,d,e} w\p{e,f},
-\end{align}
-where the proportionality is up to an appropriate symmetry factor (in
-this case, $1/2!\times1/2!$, as clarified in Section
-\ref{sec:symmetry_factors}).  Crosses, meanwhile, correspond to
-indices that are {\it only} summed over the values of ``filled dot
-indices'', e.g.
+  = \sum_{\p{a,b,e}\in\D_N\p{3}} \sum_{c,d,f\in\ZZ_N}
+  v\p{a,b,c,d,e} w\p{e,f}.
+\end{align}
+Crosses, meanwhile, correspond to indices that are {\it only} summed
+over the values of ``filled dot indices'', e.g.
 \begin{align}
   \diagram{example_x}
-  = \sum_{\p{a,e}\in\D_N\p{2}} \sum_{b,c\in\set{a,e}}
-  \sum_{d\in\ZZ_N} u\p{a,b,c} v\p{b,d,e} w\p{b,c,e}.
+  = \sum_{\p{a,e}\in\D_N\p{2}} \sum_{b,c\in\set{a,e}} \sum_{d\in\ZZ_N}
+  u\p{a,b,c} v\p{b,d,e} w\p{b,c,e}.
 \end{align}
@@
 sum over $a\in\ZZ_N\setminus\set{b,c}$ into a difference of sums over
-$a\in\ZZ_N$ and $a\in\set{b,c}$.  The sum over $a\in\ZZ_N$ to compute
-$v\p{\circ,b}$ has $O\p{N^2}$ cost, and can be performed prior to
-evaluating the diagram in \eqref{eq:example_elim_o}.  The
-decomposition in \eqref{eq:example_elim} therefore splits one
-$O\p{N^3}$ diagram into two $O\p{N^2}$ diagrams.
+$a\in\ZZ_N$ and $a\in\set{b,c}$.  Together, the sums over $a\in\ZZ_N$
+to compute all $v\p{\circ,b}$ have an $O\p{N^2}$ cost, and can be
+performed prior to evaluating the diagram in
+\eqref{eq:example_elim_o}.  The decomposition in
+\eqref{eq:example_elim} therefore splits one $O\p{N^3}$ diagram into
+two $O\p{N^2}$ diagrams.
 
@@
   = \sum_{\p{b,c}\in\D_N\p{2}} v\p{c,b} w\p{b,c}
-  = 2 \diagram{example_elim_x_full}.
+  = \diagram{example_elim_x_full},
   \label{eq:example_elim_final}
 \end{align}
-The added factor of 2 accounts for the fact that the two diagrams in
-\eqref{eq:example_elim_final} have different symmetry factors, whereas
-re-writing index constraints (i.e.~by eliminating crosses) does not
-affect any numerical prefactors on the tensor contraction represented
-by a diagram.  We can generally repeat the above process of
-sequentially eliminating filled dots and crosses until only empty dots
-remain.  The only complication in carrying out this process is that we
-have to manually keep track of symmetry factors, such as the factor of
-2 above.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Symmetry factors}
-\label{sec:symmetry_factors}
-
-As seen in \eqref{eq:example_elim_final}, the process of simplifying
-diagrams with permutationally symmetric tensors generally involves
-symmetry factors that need to be kept track of by hand.  Here we
-discuss the rules governing symmetry factors that appear when
-decomposing such diagrams.  In general, eliminating a filled dot from
-a region with $F$ filled dots makes a diagram pick up a factor of
-$1/F$, e.g.
-\begin{align}
-  \diagram{example_sym}
-  = \f14 \diagram{example_sym_o}
-  - \f14 \diagram{example_sym_x},
-\end{align}
-Similarly, adding an empty dot to a region with $E$ empty dots makes
-the diagram pick up a factor of $E+1$, so
-\begin{align}
-  \diagram{example_sym_o}
-  = \f23 \diagram{example_sym_oo}
-  - \f13 \diagram{example_sym_ox}.
-\end{align}
-Finally, eliminating a cross by moving a filled dot from an ``old''
-region with $F_{\t{old}}$ filled dots into a ``new'' region with
-$F_{\t{new}}$ filled dots picks up an overall factor of
-$F_{\t{new}}+1$, so
-\begin{align}
-  \diagram{example_sym_x} = 2 \diagram{example_sym_x_elim},
-\end{align}
-where the factor of $1/F_{\t{old}}$ that is acquired due to removing a
-dot from the old region is, loosely speaking, canceled out by the
-$F_{\t{old}}$ choices of which dot to move.  More precisely,
-eliminating a cross and moving a filled dot corresponds to forcing the
-index associated with the cross to be equal to the index associated
-with the filled dot, and there are $F_{\t{old}}$ such indices to
-choose from.
+If there are multiple filled dots for the cross to eliminate, then the
+result is a sum over all choices of filled dots:
+\begin{align}
+  \diagram{example_elim_color_start}
+  \sim
+  \diagram{example_elim_color_x}
+  = \diagram{example_elim_color_x_r}
+  + \diagram{example_elim_color_x_m}
+  + \diagram{example_elim_color_x_b}
+  \sim
+  3 \diagram{example_elim_color_end}.
+\end{align}
+We can generally repeat the above process of sequentially eliminating
+filled dots and crosses until only empty dots remain, at which point
+we have made all indices in a diagram independent.
 
@@
 
-The process of constructing and simplifying diagrams to compute
-coefficients of the diagrammatic operator product expansion in
-\eqref{eq:sym_prod_diagrams} is systematic enough to be converted into
-a computer algorithm.  The last step necessary to automate the
-evaluation of \eqref{eq:sym_prod_diagrams} is the expansion of the
-multi-spin operators $\m X\p{\ell}$ in a basis for the PS manifold
-$\M_0$.  Without loss of generality, we consider a single $M$-spin
-operator $X$.  Our task is essentially to find coefficients of the
-expansion
+The process of constructing and simplifying diagrams to compute scalar
+coefficients of the expansions in \eqref{eq:sym_prod_diagrams} and
+\eqref{eq:sum_diagrams_alt} is systematic enough to perform
+algorithmically.  The last step necessary to automate the evaluation
+of these expansions is the projection of multi-spin operators
+$\ot{\m X}\p{\ell}$ onto the PS manifold $\M_0$.  Without loss of
+generality, we consider a single $M$-spin operator $X$.  Our task is
+essentially to find coefficients of the expansion
 \begin{align}
@@
   &&
-  \C\p{a} \equiv \f{\p{\sum_\mu a_\mu}!}{\prod_\mu a_\mu!},
+  \C\p{a} \equiv \f{\p{\sum_\mu a_\mu}!}{\prod_\nu a_\nu!},
 \end{align}
@@
 with respect to transposition; in practice, one can simply sum over
-$\alpha\in\A_n\p{M}$ and set $\beta=b-a+\alpha$.  Note that, by slight
-abuse of notation, the operator $X$ on the left of
-\eqref{eq:multi_body_eval} acts on an arbitrary choice of $M$ spins
-(out of $N$), whereas the operator $X$ on the right of
-\eqref{eq:multi_body_eval} is simply an $M$-spin operator, with matrix
-elements $\bk{\alpha|X|\beta}$ evaluated with respect to the $M$-spin
-states $\ket\alpha,\ket\beta$.
+$\alpha\in\A_n\p{M}$ and set $\beta=b-a+\alpha$ (throwing out terms
+with any $\beta_\mu<0$).  Note that, by slight abuse of notation, the
+operator $X$ on the left of \eqref{eq:multi_body_eval} acts on an
+arbitrary choice of $M$ spins (out of $N$), whereas the operator $X$
+on the right of \eqref{eq:multi_body_eval} is simply an $M$-spin
+operator, with matrix elements $\bk{\alpha|X|\beta}$ evaluated with
+respect to the $M$-spin states $\ket\alpha,\ket\beta$.
 
@@
 \begin{align}
-  X_{\m v} Y_{\m w}
-  = \sp{\sum_{p,q\in\ZZ_N} v_p w_q X_p Y_q}
+  \O\p{\m v,X} \O\p{\m w,Y}
+  = \sum_{i,j} v_i w_j X_i Y_j
   \EQPS \diagram{single_body_0} X_1 Y_2
@@
 \end{align}
-where $\P_0$ is a projector onto the PS manifold $\M_0$; $p,q$ index
-individual spins; $v_p,w_p$ are scalars; and $X,Y$ are single-spin
-operators.  This product appears, for example, in the calculation of
-the second-order effective Hamiltonian $H_1^{(2)}$ induced on the
-permutationally symmetric manifold $\M_0$ by a single-body
-perturbation.  Defining
-\begin{align}
-  \v{\m m} \equiv \sum_{p\in\ZZ_N} m_p \ket{p},
-  &&
-  \col{m} \equiv \sum_{p\in\ZZ_N} m_p,
-\end{align}
-for both $\m m\in\set{\m v,\m w}$, we can simplify
-\begin{align}
-  \diagram{single_body_1} = \v{\m v}\c\v{\m w},
-\end{align}
-and
+where $i,j$ index individual spins that the single-spin operators
+$X,Y$ act on; and $\m m\equiv\sum_i m_i\ket{i}$ for
+$\m m\in\set{\m v,\m w}$ are vectors of scalar coefficients.  This
+product appears, for example, in the calculation of the second-order
+effective Hamiltonian $H_1^{(2)}$ induced on the permutationally
+symmetric manifold $\M_0$ by a single-body perturbation.  Using the
+rules that we worked out in Appendix \ref{sec:diagrams} for reducing
+diagrams, we can simplify
 \begin{align}
@@
   = \diagram{single_body_0_oo} - \diagram{single_body_1}
-  = \col{v}\,\col{w} - \v{\m v} \c\v{\m w},
-\end{align}
+  = \col{v}\,\col{w} - \m v \c \m w,
+\end{align}
+where $\col{m} \equiv \sum_i m_i$ for both $\m m\in\set{\m v,\m w}$,
 so
 \begin{align}
-  \sum_{p,q\in\ZZ_N} v_p w_q X_p Y_q
-  \EQPS \col{v}\,\col{w} X_1 Y_2
-  - \v{\m v}\c\v{\m w} \p{X_1 Y_2 - X_1 Y_1},
+  \O\p{\m v,X} \O\p{\m w,Y}
+  \EQPS \p{\col{v}\,\col{w} - \m v\c\m w} X_1 Y_2
+  + \m v\c\m w X_1 Y_1.
 \end{align}
 In order to write this result in terms of collective operators
-$\col{Z} \equiv \sum_{p\in\ZZ_N} Z_p$, we expand
+$\col{Z} \equiv \sum_i Z_i$, we expand
 \begin{align}
@@
 \begin{align}
-  \sum_{p,q\in\ZZ_N} v_p w_q X_p Y_q
-  \EQPS \f{\col{v}\,\col{w}}{N\p{N-1}}
+  \O\p{\m v,X} \O\p{\m w,Y}
+  &\EQPS \f{\col{v}\,\col{w} - \m v\c\m w}{N\p{N-1}} \,
+  \col{X}\, \col{Y}
+  - \f{\col{v}\,\col{w} - N \m v\c\m w}{N\p{N-1}} \, \col{XY} \\
+  &\EQPS \f{\col{v}\,\col{w}}{N\p{N-1}}
   \p{\col{X}\,\col{Y} - \col{XY}}
-  - \f{\v{\m v}\c\v{\m w}}{N\p{N-1}}
+  - \f{\m v\c\m w}{N\p{N-1}}
   \p{\col{X}\,\col{Y} - N\col{XY}}.
@@
 Here we simplify the two-body product
-\begin{multline}
-  X_{\m v} Y_{\m w} = \sp{\sum_{\set{k,\ell},\set{p,q}\in\C_N\p{2}}
-    v_{k\ell} w_{pq} X_{k\ell} Y_{pq}} \\
-  \EQPS \diagram{two_body_0} X_{1,2} Y_{3,4} + \diagram{two_body_1}
-  X_{1,2} Y_{2,3} + \diagram{two_body_2} X_{1,2} Y_{1,2},
-\end{multline}
-where $\P_0$ is a projector onto the PS manifold $\M_0$; $k,\ell,p,q$
-index individual spins that the two-body operators $X_{k\ell},Y_{pq}$
-act on; and $v_{k\ell},w_{rs}$ are scalar coefficients with
-$m_{pq}=m_{qp}$ and $m_{pp}=0$ for each of $\m m\in\set{\m v,\m w}$.
-This product appears, for example, in the calculation of the
-second-order effective Hamiltonian $H_2^{(2)}$ induced on the PS
-manifold $\M_0$ by a permutationally symmetric two-body perturbation.
-Defining
-\begin{align}
-  \v{\m m} \equiv \sum_{\set{p,q}\in\C_N\p{2}} m_{pq} \ket{p,q},
-  &&
-  \v m \equiv \sum_{p,q\in\ZZ_N} m_{pq} \ket{p},
-  &&
-  \col{m} \equiv \sum_{\set{p,q}\in\C_N\p{2}} m_{pq},
-\end{align}
-for both $\m m\in\set{\m v,\m w}$, we can write
-\begin{align}
-  \diagram{two_body_2} = \v{\m v} \c\v{\m w},
-\end{align}
-and
+\begin{align}
+  \O\p{\m v,X} \O\p{\m w,Y}
+  = \sum_{i\ne j} \sum_{k\ne\ell}
+  v_{ij} w_{k\ell} X_{ij} Y_{k\ell},
+\end{align}
+where $i,j,k,\ell$ index individual spins that the two-spin operators
+$X,Y$ act on; and $\m m\equiv\sum_{i,j}m_{ij}\op{i}{j}$ for
+$\m m\in\set{\m v,\m w}$ are matrices of scalar coefficients, with
+$m_{ii}=0$.  This product appears, for example, in the calculation of
+the second-order effective Hamiltonian $H_2^{(2)}$ induced on the
+permutationally symmetric manifold $\M_0$ by a two-body perturbation.
+The diagrammatic expansion of this operator is
+\begin{align}
+  \O\p{\m v,X} \O\p{\m w,Y}
+  \EQPS \diagram{two_body_0} \spp{X,Y}_4 + \diagram{two_body_1}
+  \spp{X,Y}_3 + \diagram{two_body_2} \spp{X,Y}_2,
+  \label{eq:two_prod_pair}
+\end{align}
+where $\spp{X,Y}_q$ denotes a sum over all $q$-spin operators that can
+be constructed from $X$ and $Y$, i.e.
+\begin{align}
+  \spp{X,Y}_4 \EQPS X_{1,2} Y_{3,4}, &&
+  \spp{X,Y}_2 \EQPS X_{1,2} Y_{1,2} + X_{1,2} Y_{2,1},
+\end{align}
+\begin{align}
+  \spp{X,Y}_3 \EQPS X_{1,2} Y_{1,3} + X_{1,2} Y_{3,1}
+  + X_{1,2} Y_{2,3} + X_{1,2} Y_{3,2}.
+\end{align}
+To minimize the computational cost of the diagrams in
+\eqref{eq:two_prod_pair}, we reduce
 \begin{align}
@@
   &= \sp{\diagram{two_body_1_oo} - \diagram{two_body_1_ox}}
-  - \sp{2\diagram{two_body_2}} \label{eq:mid_zero} \\
-  &= \v v \c \v w - 2 \v{\m v} \c \v{\m w},
-\end{align}
-where the middle diagram in \eqref{eq:mid_zero} contains an empty sum
-for the cross, and is therefore equal to zero.  Finally, with a bit of
-more work (or a computer) we can work out that
+  - \sp{\diagram{two_body_2}} \label{eq:mid_zero} \\
+  &= \diagram{two_body_1_ooo} - \diagram{two_body_2_oo}
+\end{align}
+where the middle diagram in \eqref{eq:mid_zero} is zero from the fact
+that $w_{ii}=0$, and
 \begin{align}
   \diagram{two_body_0}
-  = \diagram{two_body_0_oooo} - \diagram{two_body_1_ooo}
-  + \diagram{two_body_2_oo}
-  = \col{v}\,\col{w} - \v v\c\v w + \v{\m v}\c\v{\m w}.
-\end{align}
-Altogether,
-\begin{align}
-  X_{\m v} Y_{\m w}
-  \EQPS \p{\col{v}\,\col{w} - \v v\c\v w + \v{\m v}\c\v{\m w}}
-  X_{1,2} Y_{3,4}
-  + \p{\v v\c\v w - 2 \v{\m v} \c \v{\m w}} X_{1,2} Y_{2,3}
-  + \v{\m v} \c \v{\m w}\, X_{1,2} Y_{1,2}.
+  &= \sp{\diagram{two_body_0_o}} - \sp{\diagram{two_body_0_x}} \\
+  &= \sp{\diagram{two_body_0_oo} - \diagram{two_body_0_ox}}
+  - \sp{2 \diagram{two_body_1}} \\
+  &= \sp{\diagram{two_body_0_oooo} - 2 \diagram{two_body_1_o}}
+  - \sp{2 \diagram{two_body_1_ooo} - 2 \diagram{two_body_2_oo}} \\
+  &= \diagram{two_body_0_oooo} - 4 \diagram{two_body_1_ooo}
+  + 2 \diagram{two_body_2_oo}.
+\end{align}
+If the matrices $\m v,\m w$ are symmetric, then we can define
+\begin{align}
+  \v{\m m} \equiv \sum_{i\ne j} m_{ij} \ket{ij},
+  &&
+  \v m \equiv \sum_{i,j} m_{ij} \ket{i},
+  &&
+  \col{m} \equiv \sum_{i,j} m_{ij},
+\end{align}
+for both $\m m\in\set{\m v,\m w}$, and expand
+\begin{align}
+  \diagram{two_body_2} = \v{\m v} \c\v{\m w},
+  &&
+  \diagram{two_body_1} =
+  \v v \c \v w - \v{\m v} \c \v{\m w},
+  &&
+  \diagram{two_body_0}
+  = \col{v}\,\col{w} - 4 \v v\c\v w + 2 \v{\m v}\c\v{\m w}.
 \end{align}
 In the special case that $X=Y=Z\otimes Z$ with a single-spin operator
-$Z$ for which $Z^2=1$, we can further simplify
-\begin{align}
-  \sum_{\p{k,\ell},\p{p,q}\in\D_N\p{2}} v_{k\ell} w_{pq}
-  Z_k Z_\ell Z_p Z_q
-  \EQPS \p{\col{v}\,\col{w} - \v v\c\v w + \v{\m v}\c\v{\m w}}
-  Z^{\otimes 4}
-  + \p{\v v\c\v w - \v{\m v} \c \v{\m w}} Z^{\otimes 2}
-  + \v{\m v}\c\v{\m w},
-\end{align}
-which we can write in terms of collective operators
-$\col{Z}\equiv\sum_{p\in\ZZ_N}Z_p$ as
-\begin{multline}
-  \sum_{\p{k,\ell},\set{p,q}\in\C_N\p{2}} v_{k\ell} w_{pq}
-  Z_k Z_\ell Z_p Z_q \\
-  \EQPS \f{\col{v}\,\col{w} - \v v\c\v w + \v{\m v}\c\v{\m w}}{N!_4}
-  \sp{\col{Z}^4 - 2\p{3N-4} \col{Z}^2}
-  + \f{\v v\c\v w - 2 \v{\m v} \c \v{\m w}}{N!_2}\, \col{Z}^2 \\
-  + \f{3\,\col{v}\,\col{w} - N \v v\c\v w + N\p{N-2}\v{\m v}\c\v{\m
-      w}}{\p{N-1}\p{N-3}},
-\end{multline}
+$Z$ for which $Z^2=1$, we can simplify
+\begin{align}
+  \spp{Z\otimes Z,Z\otimes Z}_4 \EQPS Z^{\otimes 4},
+  &&
+  \spp{Z\otimes Z,Z\otimes Z}_3 \EQPS 4 Z^{\otimes 2},
+  &&
+  \spp{Z\otimes Z,Z\otimes Z}_2 \EQPS 2,
+\end{align}
+which implies that
+\begin{align}
+  \O\p{\m v,Z\otimes Z} \O\p{\m w,Z\otimes Z}
+  \EQPS A_4 Z^{\otimes 4} + A_2 Z^{\otimes 2} + A_0,
+\end{align}
+with
+\begin{align}
+  A_4 \equiv \diagram{two_body_0}
+  = \col{v}\,\col{w} - 4 \v v\c\v w + 2 \v{\m v}\c\v{\m w},
+  &&
+  A_2 \equiv 4 \diagram{two_body_1}
+  = 4 \v v \c \v w - 4 \v{\m v} \c \v{\m w}
+\end{align}
+\begin{align}
+  A_0 \equiv 2 \diagram{two_body_2} = 2 \v{\m v} \c\v{\m w}.
+\end{align}
+In terms of collective operators $\col{Z}\equiv\sum_iZ_i$, we have
+\begin{align}
+  \O\p{\m v,Z\otimes Z} \O\p{\m w,Z\otimes Z}
+  \EQPS \tilde A_4 \col{Z}^4 + \tilde A_2 \col{Z}^2 + \tilde A_0,
+\end{align}
 where
 \begin{align}
-  N!_k \equiv \prod_{j=0}^{k-1} \p{N-j} = \f{N!}{\p{N-k}!}
-\end{align}
-denotes a falling factorial.
+  \tilde A_4 \equiv \f{A_4}{N!_4},
+  &&
+  \tilde A_2 \equiv \f{A_2}{N!_2} - 2\p{3N-4} \f{A_4}{N!_4},
+  &&
+  \tilde A_0 \equiv A_0 - \f{A_2}{N-1} + \f{3A_4}{\p{N-1}\p{N-3}},
+\end{align}
+and
+\begin{align}
+  N!_q \equiv \sum_{j=0}^{q-1} \p{N-j}
+\end{align}
+is a falling factorial.
 
@@
 \begin{align}
-  Z_{\m u\m v\m w}
-  \equiv \sum_{\set{k,\ell},\set{p,q},\set{r,s}\in\C_N\p{2}}
-    u_{k\ell} v_{pq} w_{rs} Z_k Z_\ell Z_p Z_q Z_r Z_s,
-  &&
-  Z^2 = 1,
-\end{align}
-projected onto the PS manifold $\M_0$, where $k,\ell,p,q,r,s$ index
-individual spins; $Z$ is a single-spin operator; and
-$u_{k\ell},v_{pq},w_{rs}$ are scalar coefficients with $m_{pq}=m_{qp}$
-and $m_{pp}=0$ for each of $\m m\in\set{\m u,\m v,\m w}$.  This
-product appears, for example, in the projection of Ising-type
-($s_\z\otimes s_\z$) interactions onto the two-body excitation
-manifold $\M_2$.  Collecting terms according to the number of $Z$
-operators that remain after accounting for $Z^2=1$, we find that
-\begin{align}
-  Z_{\m u\m v\m w} \EQPS A_6 Z^{\otimes 6} + A_4 Z^{\otimes 4}
-  + A_2 Z^{\otimes 2} + A_0,
+  \O\p{\m u,\m v,\m w,Z}
+  &\equiv \O\p{\m u,Z\otimes Z} \O\p{\m v,Z\otimes Z}
+  \O\p{\m w,Z\otimes Z} \\
+  &= \sum_{i\ne j} \sum_{k\ne\ell} \sum_{p\ne q}
+  u_{ij} v_{k\ell} w_{pq} Z_i Z_j Z_k Z_\ell Z_p Z_q,
+\end{align}
+projected onto the PS manifold $\M_0$, where $i,j,k,\ell,p,q$ index
+individual spins; $Z$ is a single-spin Ising-like operator, with
+$Z^2=1$; and $\m m\equiv\sum_{i,j} m_{ij}\op{i}{j}$ for
+$\m m\in\set{\m u,\m v,\m w}$ are matrices of scalar coefficients,
+with $m_{ii}=0$.  This product appears, for example, in the projection
+of Ising interactions onto the two-body excitation manifold $\M_2$.
+Collecting terms according to the number of $Z$ operators that remain
+after accounting for $Z^2=1$, we find that
+\begin{align}
+  \O\p{\m u,\m v,\m w,Z}
+  \EQPS A_6 Z^{\otimes 6} + A_4 Z^{\otimes 4} + A_2 Z^{\otimes 2} + A_0,
   \label{eq:triple_multi}
@@
   &&
-  A_4 \equiv \diagram{triple_01} + \diagram{triple_1},
-  &&
-  A_0 \equiv \diagram{triple_0111},
+  A_4 \equiv 4 \diagram{triple_01} + 8 \diagram{triple_1},
+  &&
+  A_0 \equiv 8 \diagram{triple_0111},
   \label{eq:triple_64}
@@
 \begin{align}
-  A_2 \equiv \diagram{triple_011} + \diagram{triple_02}
-  + \diagram{triple_11} + \diagram{triple_2},
+  A_2 \equiv 8 \diagram{triple_011} + 2 \diagram{triple_02}
+  + 8 \diagram{triple_11} + 4 \diagram{triple_2},
   \label{eq:triple_2}
@@
   &&
-  D_4 \equiv \diagram{triple_01_o} - 2 \diagram{triple_1_o},
-  &&
-  D_0 \equiv \diagram{triple_0111_o},
-\end{align}
-\begin{align}
-  D_2 \equiv \diagram{triple_011_o}
-  + \diagram{triple_02_o}
-  - 2 \diagram{triple_11_o}
-  + 4 \diagram{triple_2_o}.
-\end{align}
-Defining, for each of $\m m\in\set{\m u,\m v,\m w}$,
-\begin{align}
-  \v{\m m} \equiv \sum_{\set{p,q}\in\C_N\p{2}} m_{pq} \ket{\set{p,q}},
-  &&
-  \v m \equiv \sum_{p\in\ZZ_N} m_p \ket{p},
-  &&
-  m_p \equiv \sum_{q\in\ZZ_N} m_{pq},
-  &&
-  \col{m} \equiv \sum_{\set{p,q}\in\C_N\p{2}} m_{pq},
-\end{align}
-we can expand
+  D_4 \equiv 4 \diagram{triple_01_o} - 16 \diagram{triple_1_o},
+  &&
+  D_0 \equiv 8 \diagram{triple_0111_o},
+\end{align}
+\begin{align}
+  D_2 \equiv 8 \diagram{triple_011_o}
+  + 2 \diagram{triple_02_o}
+  - 16 \diagram{triple_11_o}
+  + 16 \diagram{triple_2_o}.
+\end{align}
+If $m_{ij}=m_{ji}$ for all $\m m\in\set{\m u,\m v,\m w}$, then we can
+define
+\begin{align}
+  \v{\m m} \equiv \sum_{i\ne j} m_{ij} \ket{ij},
+  &&
+  \v m \equiv \sum_i m_i \ket{i},
+  &&
+  m_i \equiv \sum_j m_{ij},
+  &&
+  \col{m} \equiv \sum_{i,j} m_{ij},
+\end{align}
+and expand
 \begin{align}
@@
   &&
-  D_4 = \col{u}\,\v v\c\v w + \col{v}\,\v w\c\v u
-  + \col{w}\,\v u\c\v v - 2 \sum_{p\in\ZZ_N} u_p v_p w_p,
-  &&
-  D_0 = \tr\p{\m u \c \m v \c \m w},
+  D_4 = 4\p{\col{u}\,\v v\c\v w + \col{v}\,\v w\c\v u
+  + \col{w}\,\v u\c\v v} - 16 \sum_i u_i v_i w_i,
+  &&
+  D_0 = 8 \tr\p{\m u \c \m v \c \m w},
 \end{align}
 \begin{multline}
-  D_2 = \v u \c\m v\c\v w + \v v \c\m w\c\v u + \v w \c\m u\c\v v
-  + \col{u}\,\v{\m v}\c\v{\m w} + \col{v}\,\v{\m w}\c\v{\m u}
-  + \col{w}\,\v{\m u}\c\v{\m v} \\
-  - 2 \sum_{p,q\in\ZZ_N} \p{u_p v_{pq} w_{pq}
-    + v_p w_{pq} u_{pq} + w_p u_{pq} v_{pq}}
-  + 4 \sum_{\set{p,q}\in\C_N\p{2}} u_{pq} v_{pq} w_{pq}.
+  D_2 = 8\p{\v u \c\m v\c\v w + \v v \c\m w\c\v u + \v w \c\m u\c\v v}
+  + 2\p{\col{u}\,\v{\m v}\c\v{\m w} + \col{v}\,\v{\m w}\c\v{\m u}
+  + \col{w}\,\v{\m u}\c\v{\m v}} \\
+  - 16 \sum_{i,j} \p{u_i v_{ij} w_{ij}
+    + v_i w_{ij} u_{ij} + w_i u_{ij} v_{ij}}
+  + 16 \sum_{i,j} u_{ij} v_{ij} w_{ij}.
 \end{multline}
-Finally, we can also write the product $Z_{\m u\m v\m w}$ in
-\eqref{eq:triple_multi} in terms of the collective operator
-$\col{Z} \equiv \sum_{p\in\ZZ_N} Z_p$:
-\begin{align}
-  Z_{\m u\m v\m w} \EQPS
-  \tilde A_6 \col{Z}^6 + \tilde A_4 \col{Z}^4
+Finally, we can also write the product $\O\p{\m u,\m v,\m w,Z}$ in
+terms of the collective operator $\col{Z} \equiv \sum_i Z_i$:
+\begin{align}
+  \O\p{\m u,\m v,\m w,Z}
+  \EQPS \tilde A_6 \col{Z}^6 + \tilde A_4 \col{Z}^4
   + \tilde A_2 \col{Z}^2 + \tilde A_0,
-  \label{eq:triple_col}
 \end{align}
@@
 \end{align}
+and
+\begin{align}
+  N!_q \equiv \sum_{j=0}^{q-1} \p{N-j}
+\end{align}
+is a falling factorial.
 

### 2020-04-25 collective_sun/scripts/operator_product_methods.py

@@ -138,15 +138,13 @@ def reduce_diagram(diagram):
             empty_copy = diagram.copy()
             empty_copy[region] = ( markers[0]-1, markers[1], markers[2]+1 )
             empty_diag = reduce_diagram(empty_copy)
-            empty_sym = ( markers[2]+1 ) / markers[0] # symmetry factor
 
             # convert a filled dot into a cross
             cross_copy = diagram.copy()
             cross_copy[region] = ( markers[0]-1, markers[1]+1, markers[2] )
             cross_diag = reduce_diagram(cross_copy)
-            cross_sym = 1 / markers[0] # symmetry factor
 
-            return empty_sym * empty_diag - cross_sym * cross_diag
+            return empty_diag - cross_diag
 
     # otherwise, eliminate any crosses in a diagram
     else:
@@ -169,8 +167,8 @@ def reduce_diagram(diagram):
                 if joint_markers == None: joint_markers = (0,0,0)
                 new_diagram[joint_region] = ( joint_markers[0]+1, ) + joint_markers[-2:]
 
-                symmetry_factor = new_diagram[joint_region][0]
-                return symmetry_factor * reduce_diagram(new_diagram)
+                multiplicity = diagram[other_region][0]
+                return multiplicity * reduce_diagram(new_diagram)
 
             return sum([ _take_from_region(other_region)
                          for other_region, other_markers in diagram.items()
@@ -185,15 +183,6 @@ def reduce_diagram(diagram):
 # get appropriate data type to use, given a list of operators
 def _get_dtype(operators):
     return type( np.product([ operator.ravel()[0] for operator in operators ]) )
-
-# symmetrize a tensor operator under all permutations of its target spaces
-def symmetrize_operator(oper):
-    subsystems = oper.ndim//2
-    def _permuted_oper(perm):
-        return np.transpose(oper, perm + tuple( np.array(perm) + subsystems ) )
-    perm_sum = sum( _permuted_oper(perm) for perm in it.permutations(range(subsystems)) )
-    num_perms = np.math.factorial(subsystems)
-    return perm_sum / num_perms
 
 # contract operators according to a set diagram
 def contract_ops(base_ops, diagram):
@@ -228,7 +217,7 @@ def contract_ops(base_ops, diagram):
     return np.array( tf.einsum(contraction, *base_ops).numpy() )
 
 # contract tensors according to a set/index diagram
-# note: assumes translational invariance by default
+# note: assumes that all tensors are permutationally symmetric
[…]

### 2020-04-25 collective_sun/scripts/shell_model.py

@@ -33,7 +33,8 @@ zz_couplings = np.arange(-3, +4.01, 0.1)
 
 # values of the ZZ coupling to inspect more closely: half-integer values
 inspect_zz_couplings = [ zz_coupling for zz_coupling in zz_couplings
-                         if np.allclose(zz_coupling % 0.5, 0) ]
+                         if np.allclose(zz_coupling % 0.5, 0)
+                         or np.allclose(zz_coupling % 0.5, 0.5) ]
 inspect_sim_time = 2
 
 max_time = 10 # in units of J_\perp
@@ -168,7 +169,7 @@ except:
     print(runtime(), "building perturbation operator")
     sys.stdout.flush()
     shell_coupling_op \
-        = build_shell_operator([sunc["mat"]], [local_ops["ZZ"]], sunc, sunc["TI"])
+        = build_shell_operator([sunc["mat"]], [local_ops["ZZ"]/4], sunc, sunc["TI"])
     shell_coupling_op = np.einsum("ikjk->ijk", shell_coupling_op)
     np.savetxt(coupling_file, shell_coupling_op.flatten())
 
@@ -251,8 +252,7 @@ def coherent_spin_state(vec):
 # simulate!
 ##########################################################################################
 
-# note: extra factor of 1/2 in che_eff_bare for compatibility with Chunlei's work
-chi_eff_bare = 1/4 * sunc["mat"].sum() / np.math.comb(spin_num,2)
+chi_eff_bare = sunc["mat"].sum() / (spin_num * (spin_num-1))
 state_X = coherent_spin_state([0,1,0])
 
 def _states(initial_state, zz_sun_ratio, times):
@@ -276,8 +276,7 @@ def simulate(zz_coupling, sim_time = None, max_tau = 2, points = 500):
     times = np.linspace(0, sim_time, points)
 
     # compute states at all times of interest
-    # note: factor of 1/2 included for compatibility with Chunlei's work
-    states = _states(state_X, zz_sun_ratio/2, times)
+    states = _states(state_X, zz_sun_ratio, times)
 
     # compute collective spin correlators
     def val(mat, state):

### 2020-04-25 collective_sun/scripts/simplify_product_diagrams.py

@@ -7,34 +7,37 @@ import copy
 from itertools_extension import set_diagrams
 from operator_product_methods import diagram_vec
 
-dimensions = [ 2, 2, 2 ]
+dimensions = [ 2, 2 ]
 
-# construct a random symmetric tensor with zeros on all diagonal blocks
-def random_tensor(dimension):
-    tensor = np.zeros((spins,)*dimension)
-    for comb in it.combinations(range(spins), dimension):
-        num = np.random.rand()
-        for perm in it.permutations(comb):
-            tensor[perm] = num
-    return tensor
+def num_assignments(diagram):
+    factors = set.union(*[ set(subset) for subset in diagram.keys() ])
+    dimensions = { factor : sum( points for subset, points in diagram.items()
+                                 if factor in subset )
+                   for factor in factors }
+    num = np.prod([ np.math.factorial(dimension) for dimension in dimensions.values() ])
+    den = np.prod([ np.math.factorial(points) for points in diagram.values() ])
+    return num / den
+
+def diagram_weight(diagram):
+    return sum( num_indices
+                for ops, num_indices in diagram.items()
+                if len(ops) % 2 == 1 )
 
 # collect diagrams by weight of the multi-body operator they are associated with
 weight_vec = {}
 for diagram in set_diagrams(dimensions):
-    weight = sum( num_indices
-                  for ops, num_indices in diagram.items()
-                  if len(ops) % 2 == 1 )
-    if weight not in weight_vec.keys():
-        weight_vec[weight] = diagram_vec(diagram)
-    else:
-        weight_vec[weight] += diagram_vec(diagram)
+    weight = diagram_weight(diagram)
+    vec = num_assignments(diagram) * diagram_vec(diagram)
+    if weight not in weight_vec:
+        weight_vec[weight] = 0
+    weight_vec[weight] += vec
 
 for weight, vec in weight_vec.items():
     print(weight)
     print("-"*10)
 
     def _name(diag):
-        diag_name = ""
+        diag_name = f"{diagram_weight(diag)}_"
         if (0,1,2) in diag.keys():
             diag_name += str(diag[0,1,2])
         else:
@@ -48,7 +51,7 @@ for weight, vec in weight_vec.items():
     reduced_vec = vec.reduce().join_permutations()
     names_diags_coefs = sorted(zip(map(_name, reduced_vec.diags),
                                    reduced_vec.diags,
-                                   reduced_vec.coefs))[::-1]
[…]

### 2020-04-28 collective_sun/multilevel_spin_notes.tex
changed some notation and added 'selection' rules for products of drive operators

@@
 \DeclareMathOperator{\var}{var}
+\DeclareMathOperator{\even}{even}
 
@@
   D_{LM} \equiv \f{\eta_M}{\sqrt{2}}
-  \sp{T_{L\abs{M}} + \sign\p{M} T_{L\abs{M}}^\dag},
+  \sp{T_{LM} + \sign\p{M} T_{LM}^\dag},
   &&
@@
     \sqrt{2} & M = 0 \\
-    \p{-1}^M \sign\p{M}^{3/2} & M \ne 0
+    \p{-1}^M \sign\p{M}^{M+1/2} & M \ne 0
   \end{cases},
@@
 system by $\H_n$, denote its $N$-fold tensor product by
-$\H_n^{\otimes N}$, and identify the set $\Pi\equiv\set{\Pi_\sigma}$
-of operators that permute tensor factors of $\H_n^{\otimes N}$
-according to elements $\sigma\in\SS_N$ of the symmetric group of order
-$N$.  We then expand
+$\H_n^{\otimes N}$, and identify the set $\Pi\equiv\set{\Pi_s}$ of
+operators that permute tensor factors of $\H_n^{\otimes N}$ according
+to elements $s\in\SS_N$ of the symmetric group of order $N$.  We then
+expand
 \begin{align}
@@
   T_{\ell_1 m_1} T_{\ell_2 m_2}
-  = \sum_{L,M} g_{\ell_1 m_1;\ell_2 m_2}^{LM} T_{LM},
+  = \sum_{L,M} f_{\ell_1 m_1;\ell_2 m_2}^{LM} T_{LM},
 \end{align}
@@
 \begin{align}
-  g_{\ell_1 m_1;\ell_2 m_2}^{LM}
+  f_{\ell_1 m_1;\ell_2 m_2}^{LM}
   \equiv \obk{T_{LM} | T_{\ell_1 m_1} T_{\ell_2 m_2}}
   = \tr\p{T_{LM}^\dag T_{\ell_1 m_1} T_{\ell_2 m_2}}.
-  \label{eq:trans_struct}
+  \label{eq:trans_struct_start}
 \end{align}
@@
 \end{align}
-Substituting this result into \eqref{eq:trans_struct} and expanding
-Clebsch-Gordan coefficients in terms of Wigner 3-$j$ symbols as
+Substituting this result into \eqref{eq:trans_struct_start} and
+expanding Clebsch-Gordan coefficients in terms of Wigner 3-$j$ symbols
+as
 \begin{align}
@@
 \begin{multline}
-  g_{\ell_1 m_1;\ell_2 m_2}^{LM}
+  f_{\ell_1 m_1;\ell_2 m_2}^{LM}
   = \delta_{M,m_1+m_2}
@@
 \begin{align}
-  g_{\ell_1 m_1;\ell_2 m_2}^{LM}
+  f_{\ell_1 m_1;\ell_2 m_2}^{LM}
   &= \delta_{M,m_1+m_2} \p{-1}^{2I+M}
@@
   \end{Bmatrix}.
-  \label{eq:trans_prod}
+  \label{eq:trans_struct}
 \end{align}
@@
   \bk{\ell_1 m_1; \ell_2 m_2| \ell_3 m_3}
-  = \p{-1}^{\ell_1+\ell_2-\ell_3}
+  &= \p{-1}^{\ell_1+\ell_2-\ell_3}
+  \bk{\ell_1,-m_1; \ell_2, -m_2| \ell_3 m_3} \\
+  &= \p{-1}^{\ell_1+\ell_2-\ell_3}
   \bk{\ell_2 m_2; \ell_1 m_1| \ell_3 m_3},
@@
 \begin{align}
-  g_{\ell_1 m_1;\ell_2 m_2}^{LM}
-  = \p{-1}^{\ell_1+\ell_2-L} g_{\ell_2 m_2;\ell_1 m_1}^{LM}.
+  f_{\ell_1 m_1;\ell_2 m_2}^{LM}
+  &= \p{-1}^{\ell_1+\ell_2-L} f_{\ell_1,-m_2;\ell_2,-m_2}^{L,-M} \\
+  &= \p{-1}^{\ell_1+\ell_2-L} f_{\ell_2 m_2;\ell_1 m_1}^{LM}.
 \end{align}
@@
   D_{LM} \equiv \f{\eta_M}{\sqrt{2}}
-  \sp{T_{L\abs{M}} + \sign\p{M} T_{L\abs{M}}^\dag},
+  \sp{T_{LM} + \sign\p{M} T_{LM}^\dag},
   &&
   \eta_M \equiv
-  \begin{cases}
-    \sqrt{2} & M = 0 \\
-    \p{-1}^M \sign\p{M}^{3/2} & M \ne 0
-  \end{cases},
-\end{align}
-or, equivalently,
-\begin{align}
-  D_{LM} \equiv \f{\tau_M}{\sqrt{2}}
-  \sp{T_{LM} + \sign\p{M} T_{LM}^\dag},
-  &&
-  \tau_M \equiv
   \begin{cases}
@@
 \end{align}
+or, equivalently,
+\begin{align}
+  D_{LM} \equiv \f{\eta_M}{\sqrt{2}}
+  \sum_{s\in\set{\pm1}} \epsilon_{Ms} T_{L,sM},
+  &&
+  \epsilon_{Ms} \equiv
+  \begin{cases}
+    1 & s = +1 \\
+    \p{-1}^M \sign\p{M} & s = -1
+  \end{cases},
+\end{align}
 which implies that
 \begin{align}
-  T_{LM} = \f{\tilde\tau_M}{\sqrt{2}} \p{D_{LM} + i D_{L,-M}},
-  &&
-  \tilde\tau_M \equiv
+  T_{LM} = \f{\tau_M}{\sqrt{2}}
+  \sum_{s\in\set{\pm1}} \sqrt{s} D_{L,sM},
+  &&
+  \tau_M \equiv
   \begin{cases}
     \sqrt{2}/\p{1+i} & M = 0 \\
-    \tau_M^* & M \ne 0
+    \eta_M^* & M \ne 0
   \end{cases},
 \end{align}
-where $\tau_M^*$ is the complex conjugate of $\tau_M$.  In order to
+where $\eta_M^*$ is the complex conjugate of $\eta_M$.  In order to
 evaluate the product of two drive operators, we can thus expand them
@@
 of transition operators (see Appendix \ref{sec:trans_prod}), and
-finally expand the result in terms of drive operators.  The general
-product of two drive operators takes the form
+finally expand the result in terms of drive operators.  The product of
+two drive operators is thus
 \begin{align}
   D_{\ell_1 m_1} D_{\ell_2 m_2}
-  = \f{\tau_{m_1}\tau_{m_2}}{\sqrt{2}}
-  \sum_L \p{g_{\ell_1 m_1;\ell_2 m_2}^{L,m_1+m_2}
-    E_{\ell_1 m_1;\ell_2 m_2}^{(L,+)}
-    + g_{\ell_1 m_1;\ell_2,-m_2}^{L,m_1-m_2}
-    E_{\ell_1 m_1;\ell_2 m_2}^{(L,-)}},
-\end{align}
-where in terms of
-\begin{align}
-  \epsilon_M \equiv \p{-1}^M \sign\p{M},
-\end{align}
-we have
-\begin{align}
-  E_{\ell_1 m_1;\ell_2 m_2}^{(L,+)}
-  &\equiv \sum_{s\in\set{\pm1}}
-  \f12 \sp{\tilde\tau_{m_1+m_2}
-    + i s \p{-1}^{L+\ell_1+\ell_2}
-    \epsilon_{m_1} \epsilon_{m_2} \tilde\tau_{-\p{m_1+m_2}}}
-  \sqrt{s} D_{L,s\p{m_1+m_2}},
-  \\
-  E_{\ell_1 m_1;\ell_2 m_2}^{(L,-)}
-  &\equiv \sum_{s\in\set{\pm1}}
-  \f12 \sp{\epsilon_{m_2} \tilde\tau_{m_1-m_2}
-    + i s \p{-1}^{L+\ell_1+\ell_2}
-    \epsilon_{m_1} \tilde\tau_{-\p{m_1-m_2}}}
-  \sqrt{s} D_{L,s\p{m_1-m_2}}.
-\end{align}
-If $m_1\pm m_2=0$, these operators simplify to
-\begin{align}
-  E_{\ell_1 m;\ell_2,-m}^{(L,+)}
-  &= \f1{\sqrt{2}} \sp{1 - \p{-1}^{L+\ell_1+\ell_2}
-    \p{1-\delta_{m,0}}} D_{L,0}, \\
-  E_{\ell_1 m;\ell_2 m}^{(L,-)}
-  &= \f{\epsilon_m}{\sqrt{2}}
-  \sp{1 + \p{-1}^{L+\ell_1+\ell_2}} D_{L,0},
-\end{align}
-and are otherwise
-\begin{align}
-  E_{\ell_1 m_1;\ell_2 m_2}^{(L,+)}
-  &\stackrel{m_1+m_2\ne0}{=}
-  \tau_{m_1+m_2}^* \sum_{s\in\set{\pm1}}
-  \f12 \sp{1 + s \p{-1}^{L+\ell_1+\ell_2}
-    \epsilon_{m_1} \epsilon_{m_2} \epsilon_{m_1+m_2}}
-  \sqrt{s} D_{L,s\p{m_1+m_2}}, \\
-  E_{\ell_1 m_1;\ell_2 m_2}^{(L,-)}
-  &\stackrel{m_1-m_2\ne0}{=}
-  \tau_{m_1-m_2}^* \sum_{s\in\set{\pm1}}
-  \f12 \sp{\epsilon_{m_2} + s \p{-1}^{L+\ell_1+\ell_2}
-    \epsilon_{m_1} \epsilon_{m_1-m_2}}
-  \sqrt{s} D_{L,s\p{m_1-m_2}}.
-\end{align}
-If $\p{m_1,m_2}=\p{0,m}$, then we can use the fact that
-$i\tau_m\tau_{-m}^*=\epsilon_m$ for $m\ne0$ to simplify
-\begin{align}
-  D_{\ell_1,0} D_{\ell_2 m}
-  = \sum_L g_{\ell_1,0,\ell_2,m}^{Lm}
-  \sum_{s\in\set{\pm1}} \f12 \sp{1 + s \p{-1}^{L+\ell_1+\ell_2}}
-  \sqrt{s} D_{L,sm}.
-\end{align}
+  = \sum_{L,M} g_{\ell_1m_1;\ell_2m_2}^{LM} D_{LM}
+  \label{eq:drive_prod}
+\end{align}
+where
+\begin{align}
+  g_{\ell_1m_1;\ell_2m_2}^{LM}
+  &\equiv \f{\eta_{m_1}\eta_{m_2}}{2\sqrt2}
+  \sum_{r,s_1,s_2\in\set{\pm1}} \epsilon_{m_1s_1} \epsilon_{m_2s_2}
+  \times f_{\ell_1,s_1m_1;\ell_2,s_2m_2}^{L,rM}
+  \times \tau_{s_1m_1+s_2m_2} \times \sqrt{r}.
+\end{align}
+Redefining $\p{s_1,s_2}\to\p{rs_1,rs_2}$ and using the facts that
+\begin{align}
+  f_{\ell_1,rs_1m_1;\ell_2,rs_2m_2}^{L,rM}
+  = r^{\ell_1+\ell_2+L} f_{\ell_1,s_1m_1;\ell_2,s_2m_2}^{LM},
+  &&
+  \tau_{rM} \sqrt{r} &= \tau_M \, \sigma_{Mr} \, r^M,
+\end{align}
+with
+\begin{align}
+  \sigma_{Mr} \equiv
+  \begin{cases}
+    \sqrt{r} & M = 0 \\
+    \sign\p{M}^{\frac{1-r}{2}} & M \ne 0
+  \end{cases},
+  &&
+  0^0 \equiv 1,
+\end{align}
+we can simplify
+\begin{align}
+  g_{\ell_1m_1;\ell_2m_2}^{LM}
+  = \f{\eta_{m_1}\eta_{m_2}\tau_M}{\sqrt2}
+  \sum_{s_1,s_2\in\set{\pm1}} f_{\ell_1,s_1m_1;\ell_2,s_2m_2}^{LM}
+  \times \f12 \sum_{r\in\set{\pm1}}
+  \epsilon_{m_1,r s_1} \epsilon_{m_2,r s_2} \,
+  \sigma_{Mr} \, r^{\ell_1+\ell_2+L+M}.
+  \label{eq:drive_struct}
+\end{align}
+In the case that $m_1=0$, we can simplify
+\begin{align}
+  g_{\ell_1,0;\ell_2,m}^{LM}
+  \stackrel{m\ne 0}{=} f_{\ell_1,0;\ell_2 m}^{Lm}
+  \times \delta_{\abs{M},\abs{m}} \times
+  \begin{cases}
+    \sp{\sp{\ell_1+\ell_2+L~\t{is even}}} & M = +m, \\
+    i \sp{\sp{\ell_1+\ell_2+L~\t{is odd}}} & M = -m,
+  \end{cases}
+\end{align}
+\begin{align}
+  g_{\ell_1,0;\ell_2,0}^{LM} = f_{\ell_1,0;\ell_2,0}^{LM}.
+\end{align}
+Here $\sp{\sp{X}}\equiv1$ if $X$ is true and $0$ otherwise.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Finding drive operator sub-algebras}
+
+Here we identify conditions under which the drive operator structure
+constants $g_{\ell_1m_1;\ell_2m_2}^{LM}$ vanish, which will help us
+identify subsets of all drive operators that are relevant in a
+particular physical scenario.  First, we note that
+$g_{\ell_1m_1;\ell_2m_2}^{LM}$ is guaranteed to be zero unless
+\begin{align}
+  M = s_1 m_1 + s_2 m_2 ~~ \t{for some} ~~ s_1,s_2\in\set{\pm1}.
+\end{align}
+Even if this condition is satisfied, the structure constant
+$g_{\ell_1m_1;\ell_2m_2}^{LM}$ can still vanish if the two terms in
+the sum over $r$ in \eqref{eq:drive_struct} have opposite sign.  For
+brevity, we define
+\begin{align}
+  z_{\ell_1m_1;\ell_2m_2}^{LM;s_1s_2r}
+  \equiv \epsilon_{m_1,r s_1} \epsilon_{m_2,r s_2} \,
+  \sigma_{Mr} \, r^{\ell_1+\ell_2+L+M}.
+\end{align}
+If $M=0$, then $\p{\sigma_{M,+1},\sigma_{M,-1}}=\p{1,i}$ and the
+remaining factors in $z_{\ell_1m_1;\ell_2m_2}^{LM;s_1s_2r}$ are real,
+so the sum over $r$ cannot vanish.  For the case of $M\ne0$, we can
+write
+\begin{align}
+  \epsilon_{ms} = \p{-1}^{\frac{1-s}{2}m} \sign\p{m}^{\frac{1-s}{2}},
+  &&
+  0^0 \equiv 1,
+\end{align}
+which allows us to expand
+\begin{align}
+  z_{\ell_1m_1;\ell_2m_2}^{LM;s_1s_2r}
+  \stackrel{M\ne0}{=} \p{-1}^{\frac{m_1+m_2}{2}-r\frac{M}{2}}
+  \sign\p{m_1}^{\frac{1-rs_1}{2}} \sign\p{m_2}^{\frac{1-rs_2}{2}}
+  \sign\p{M}^{\frac{1-r}{2}} r^{\ell_1+\ell_2+L+M}.
+\end{align}
+We then consider the dependence of each factor in this expansion on
+$r$:
+\begin{enumerate}
+\item $\p{-1}^{\frac{m_1+m_2}{2}-r\frac{M}{2}}$ changes sign with $r$
+  if and only if $M$ is odd,
+\item $\sign\p{m_1}^{\frac{1-rs_1}{2}}$ changes sign with $r$ if and
+  only if $m_1<0$,
+\item $\sign\p{m_2}^{\frac{1-rs_2}{2}}$ changes sign with $r$ if and
+  only if $m_2<0$,
+\item $\sign\p{M}^{\frac{1-r}{2}}$ changes sign with $r$ if and only
+  if $M<0$, and finally
+\item $r^{\ell_1+\ell_2+L+M}$ changes sign with $r$ if and only if
+  $\ell_1+\ell_2+L+M$ is odd.
+\end{enumerate}
+Altogether, the structure factor $g_{\ell_1m_1;\ell_2m_2}^{LM}$ is
+guaranteed to be zero unless an even number of the following six
+conditions are satisfied:
+\begin{align}
+  \ell_1 ~ \t{odd},
+  &&
+  \ell_2 ~ \t{odd},
+  &&
+  L ~ \t{odd}, \label{eq:L_conds} \\
+  m_1 < 0,
+  &&
+  m_2 < 0,
+  &&
+  M < 0. \label{eq:M_conds}
+\end{align}
+We now consider the structure factors
+$g_{\ell_1m_1;\ell_2m_2}^{LM,\pm}$ for (anti-)commutators of drive
+operators:
+\begin{align}
+  \sp{D_{\ell_1 m_1}, D_{\ell_2 m_2}}_\pm
+  = \sum_{L,M} g_{\ell_1m_1;\ell_2m_2}^{LM,\pm} D_{LM}.
+\end{align}
+The drive operators all self-adjoint, which implies that
+$g_{\ell_1m_1;\ell_2m_2}^{LM,\pm}$ must be purely real ($+$) or
+imaginary ($-$).  As a consequence, $g_{\ell_1m_1;\ell_2m_2}^{LM,\pm}$
+is {\it zero} unless $g_{\ell_1m_1;\ell_2m_2}^{LM}$ is purely real
+($+$) or imaginary ($-$).  These conditions under which
+$g_{\ell_1m_1;\ell_2m_2}^{LM}$ is purely real or imaginary turn out to
+be easy to determine by examination of the general product in
+\eqref{eq:drive_prod}, together with the fact that the drive operator
+$D_{LM}$ is purely real (imaginary) for $M\ge0$ ($M<0$).  It follows
+that $g_{\ell_1m_1;\ell_2m_2}^{LM}$ must be purely real (imaginary) if
+an even (odd) number of $m_1,m_2,M$ are negative.  Altogether,
+$g_{\ell_1m_1;\ell_2m_2}^{LM,\pm}$ can only be nonzero for an even
+($+$) or odd ($-$) number of $m_1,m_2,M$ are negative, which implies
+that an even ($+$) or odd ($-$) number of conditions in {\it each} of
+\eqref{eq:L_conds} and \eqref{eq:M_conds} must be satisfied in order
+to have $g_{\ell_1m_1;\ell_2m_2}^{LM,\pm}\ne0$.
+
+\red{todo: finish section}
+
+\vspace{2cm}
 
@@
 $q$; and $g_{JKL} \equiv g_{J,0;K,0}^{L,0}$ is a structure constant of
-the transition operator algebra, defined in \eqref{eq:trans_prod}.
+the transition operator algebra, defined in \eqref{eq:trans_struct}.
 Substituting the form of $B_L^{(\phi,q)}$ in \eqref{eq:B_L_phi}, we

### 2020-04-30 collective_sun/sun_perturbation.tex
minor cleanup of some text

@@
 
-\let\var\relax
-\DeclareMathOperator{\var}{var}
-\DeclareMathOperator{\cov}{cov}
 \DeclareMathOperator{\diag}{diag}
@@
 SU($n$) symmetry breaking through second order in perturbation theory,
-we develop a diagrammatic method to systematically compute all matrix
-elements of arbitrary spin operators restricted manifolds of weakly
-broken permutation symmetry, thereby enabling the numerical simulation
-of dynamics induced by non-perturbative modifications to the SU($n$)
+we develop a diagrammatic method to systematically compute matrix
+elements of arbitrary spin operators within manifolds of weakly broken
+permutational symmetry, thereby enabling the numerical simulation of
+dynamics induced by non-perturbative modifications to the SU($n$)
 Heisenberg model.  While the latter simulations generally accumulate
@@
   \caption{The permutation operator $\Pi_{ij}$ swaps the states of
-    spins $i$ and $j$, which in the case of SU(2) can be represented
-    by magnetization vectors in 3-dimensional space.  In the case of
-    SU($n$), these vectors live in a higher-dimensional space without
-    a clear geometric interpretation, but the interpretation of the
-    permutation operator $\Pi_{ij}$ remains the same.}
+    interacting spins $i$ and $j$, which in the case of SU(2) can be
+    represented by magnetization vectors in 3-dimensional space.  In
+    the case of SU($n$), these vectors live in a higher-dimensional
+    space without a clear geometric interpretation, but the
+    interpretation of the permutation operator $\Pi_{ij}$ remains the
+    same.}
   \label{fig:sketch_spins}
@@
 which gives rise to an OAT Hamiltonian at second order in perturbation
-theory, with a squeezing strength $\chi$ that is proportional to the
+theory, with a squeezing strength $\chi$ that is related to the
 variance of the magnetic field strength $B_i$.
 
-In order to go beyond first-order treatment of two-body interactions
-such as \eqref{eq:Ising}, or beyond second order in the treatment of
-single-body terms such as \eqref{eq:magnetic}, it will turn out to be
-necessary to introduce $M$-body operators with $M>2$.  Generalizing
-the above examples to the case of SU($n$) systems, we therefore
-consider a general $M$-body operator of the form
+Similarly to single- and two-body operators such as $H_{\t{B}}$ and
+$H_{\t{Ising}}$ in \eqref{eq:magnetic} and \eqref{eq:Ising}, it will
+be necessary for our purposes to additionally consider the
+introduction of $M$-body operators with $M>2$ \red{[todo: explain why
+  better]}. Generalizing the above examples to the case of SU($n$)
+systems, we therefore consider a general $M$-body operator of the form
 \begin{align}
@@
 to identify for a consistent definition of $X_k$ as an $M$-body
-operator.  In this notation, we can write the magnetic field and
+operator.  In this notation, we can write the magnetic field and Ising
 Hamiltonians in \eqref{eq:magnetic} and \eqref{eq:Ising} as
@@
 If the addition $\O\p{\m w,X}$ to the SU($n$)-symmetric Hamiltonian
-$H_0$ in \eqref{eq:H_0} is sufficiently small, then we can treat the
-effect of $\O\p{\m w,X}$ on the PS manifold $\M_0$ perturbatively.
-The effective Hamiltonians $H_{\t{eff}}^{(1)}$ and $H_{\t{eff}}^{(2)}$
-induced by $\O\p{\m w,X}$ on $\M_0$ at first and second order in
-perturbation theory are\cite{bravyi2011schrieffer}
+$H_0$ in \eqref{eq:H_0} is sufficiently small, namely with operator
+norm $\norm{\O\p{\m w,X}}$ less than half the spectral gap
+$\Delta_{\t{gap}}$ of $H_0$, $\norm{\O\p{\m w,X}}<\Delta_{\t{gap}}/2$,
+then we can treat the effect of $\O\p{\m w,X}$ on the ground-state PS
+manifold $\M_0$ perturbatively.  The effective Hamiltonians
+$H_{\t{eff}}^{(1)}$ and $H_{\t{eff}}^{(2)}$ induced by $\O\p{\m w,X}$
+on $\M_0$ at first and second order in perturbation theory
+are\cite{bravyi2011schrieffer}
 \begin{align}
@@
 where, thinking of the tensor $\m w$ as a $\abs{\D_N\p{M}}$-component
-vector, the tensor $\m w_\Delta$ can be found found by diagonalizing a
+vector, the tensor $\m w_\Delta$ can be found by diagonalizing a
 matrix $\hat{\m h}$ of dimensions
@@
   H_{\t{eff}}^{(2)}
-  = \sum_{\Delta\ne0} \f{\var\p{\m w_\Delta}}{\p{N-1}\Delta}
-  \p{\col{X}^2 - N \col{X^2}},
+  = \sum_{\Delta\ne0} \f{\m w_\Delta\c\m w_\Delta}{N\p{N-1}\Delta}
+  \p{\col{X}^2 - N \col{X^2}}.
   \label{eq:H_eff_2_1}
 \end{align}
-where
-\begin{align}
-  \var\p{\m w} \equiv \f1{\abs{\D_N\p{M}}}
-  \sum_{k\in\D_N\p{M}} \p{w_k - \mean w}^2
-\end{align}
-is the variance of $\m w$.  In the SU(2) case, this result implies
-that an inhomogeneous magnetic field ($X=s_\z$) yields the OAT
-Hamiltonian $H_{\t{OAT}}$ in \eqref{eq:OAT} at second order
-perturbation theory.
+In the SU(2) case, this result implies that an inhomogeneous magnetic
+field ($X=s_\z$) yields the OAT Hamiltonian $H_{\t{OAT}}$ in
+\eqref{eq:OAT} at second order perturbation theory.
 
@@
 \label{sec:spectral_gap}
-% todo: make a sub subsection of B.1
+
+\red{[todo: make this a subsection of Appendix
+  \ref{sec:single_body_eigenstates}]}
 
@@
 More generally, even if the couplings $\m w$ of $\O\p{\m w,X}$ do not
-obey the spatial symmetries as $H_0$, these symmetries generally endow
+obey the spatial symmetries as $H_0$, these symmetries still endow
 $\hat{\m h}$ with a block-diagonal structure that reduces the
@@
 \eqref{eq:sum_assignments} and \eqref{eq:sum_diagrams} was arbitrary,
-and we could just as well have chosen to perform an average (sum) for
-the scalar (operator) content.  For later convenience, we therefore
-define the {\it mixed} sums and averages
+and we could just as well have chosen to interchange sums and
+averages.  To simplify the diagrammatic formalism we develop in the
+following section, we therefore define the {\it mixed} sums and
+averages
 \begin{align}
@@
   \EQPS \sum_{\ell\in\L\p{g}} \m X\p{k,\ell}
-  ~\t{for any}~ k\in\D_N\p{\abs{g}}.
-\end{align}
-If all tensors $w_\alpha$ in $\m w$ (or all operators $X_\alpha$ in
-$\m X$) are permutationally symmetric, then
+  ~\t{for any}~ k\in\D_N\p{\abs{g}},
+\end{align}
+where $\ut{\m w}\p{g}$ sums over index values $k\in\D_N\p{\abs{g}}$,
+but averages over index assignments $\ell\in\L\p{g}$; and vice versa
+for $\ot{\m X}\p{g}$.  If all tensors $w_\alpha$ in $\m w$ (or all
+operators $X_\alpha$ in $\m X$) are permutationally symmetric, then
 \begin{align}

### 2020-06-09 collective_sun/sun_perturbation.tex
simplified some text, added citation to XXZ paper, added plots of two-body excitation matrices

@@
 
-\DeclareMathOperator{\diag}{diag}
+\newcommand{\oper}{\operatorname}
 
@@
 SU($n$) symmetries makes it all the more exciting that they appear in
-experimental platforms with exquisite degrees of microscopic control,
-thereby enabling the detailed study of these symmetries' consequences
-for fundamental questions in physics, as well as their practical use
-in technological applications.  SU(2)-symmetric spin interactions, for
-example, can play an important role in the development of quantum
-sensors that make use of entanglement to surpass classical limits on
-measurement precision\cite{he2019engineering}\red{[cite XXZ squeezing
-  paper]}.  The exciting prospect of similarly exploiting larger
-SU($n$) symmetries to achieve a technological advantage is still an
-unexplored avenue of research with enormous potential.
+experimental platforms with exquisite degrees of microscopic control.
+This control enables detailed studies of these symmetries'
+consequences for fundamental questions in physics, as well as their
+practical use in technological applications.  SU(2)-symmetric spin
+interactions, for example, can play an important role in the
+development of quantum sensors that make use of entanglement to
+surpass classical limits on measurement
+precision\cite{he2019engineering, perlin2020spin}.
+The exciting prospect of similarly exploiting larger SU($n$)
+symmetries to achieve a technological advantage is still an unexplored
+avenue of research with enormous potential.
+
+\red{[Paragraph on collective spin systems]}
 
@@
 theory of gapped SU($n$) ferromagnets with weakly broken SU($n$)
-symmetry, e.g.~by the presence of external fields or additional
-interaction terms.  The restriction of our theory to ferromagnetic
-interactions can be relaxed to the assumption of
+symmetry (e.g.~due to external fields or additional interaction
+terms).  The restriction of our theory to ferromagnetic interactions
+can be relaxed to the assumption of
 \begin{enumerate*}
@@
 dynamics induced by non-perturbative modifications to the SU($n$)
-Heisenberg model.  While the latter simulations generally accumulate
-uncontrolled truncation errors, we identify criteria under which
-numerical results correctly capture the qualitative dynamical behavior
-of a system.
+Heisenberg model.  While these simulations generally accumulate
+uncontrolled errors from the truncation of Hilbert space, we identify
+criteria under which numerical results correctly capture the
+qualitative dynamical behavior of a system.
 
@@
 Heisenberg model, finding a convenient way to express its Hamiltonian
-in terms of SWAP operations that are agnostic to the value of $n$.  In
-Section \ref{sec:breaking_sun}, we consider modifying the Hamiltonian
-of the bare Heisenberg model by the addition of spin operators,
-corresponding to e.g.~external fields or additional interactions, that
-weakly break SU($n$) symmetry.  We sketch out a derivation of the
-effective Hamiltonian induced by these terms through second order in
-perturbation theory, deferring more detailed calculations to
-Appendices \ref{sec:eigenstates} and \ref{sec:operator_product}.
-Finally, we discuss a numerical method to simulate dynamics within
-manifolds of weak permutational symmetry breaking in Section
-\ref{sec:shell_model}, before concluding in Section
-\ref{sec:conclusions}.  Wherever possible, we will fall back to the
-broadly familiar case of SU(2) in order to facilitate an understanding
-of the generalizations to SU($n$).
+in terms of permutation (or SWAP) operations that are agnostic to the
+value of $n$.  In Section \ref{sec:breaking_sun}, we consider
+modifying the Hamiltonian of the bare Heisenberg model by the addition
+of spin operators, corresponding to e.g.~external fields or additional
+interactions, that weakly break SU($n$) symmetry.  We sketch out a
+derivation of the effective Hamiltonian induced by these terms through
+second order in perturbation theory, deferring more detailed
+calculations to appendices.  Finally, we discuss a numerical method to
+simulate dynamics within manifolds of weak permutational symmetry
+breaking in Section \ref{sec:shell_model}, before concluding in
+Section \ref{sec:conclusions}.  Wherever possible, we will fall back
+to the broadly familiar case of SU(2) in order to facilitate an
+understanding of the generalizations to SU($n$).
 
@@
 surpass classical limits on measurement
-precision\cite{he2019engineering}\red{[cite XXZ squeezing paper]}.
+precision\cite{he2019engineering, perlin2020spin}.
 Starting with a general (gapped) SU(2)-symmetric Hamiltonian of the
@@
 be necessary for our purposes to additionally consider the
-introduction of $M$-body operators with $M>2$ \red{[todo: explain why
-  better]}. Generalizing the above examples to the case of SU($n$)
-systems, we therefore consider a general $M$-body operator of the form
+introduction of $M$-body operators with $M>2$. Generalizing the above
+examples to the case of SU($n$) systems, we therefore consider a
+general $M$-body operator of the form
 \begin{align}
@@
   \D_N\p{M} \equiv
-  \set{ k \in \ZZ_N^M : \t{all entries $k_i$ of $k$ are distinct } },
+  \set{ k \in \ZZ_N^M : \t{all entries $k_i$ of $k$ are distinct} },
   \label{eq:off_diags}
@@
   \col X \equiv \sum_{k\in\D_N\p{M}} X_k,
+  \label{eq:mean_col}
 \end{align}
@@
 $\ket\psi\in\M_0$.  The SU($n$) symmetry of $H_0$ enables such a
-decomposition of the form
+decomposition to take the form
 \begin{align}
@@
   H_{\t{eff}}^{(2)}
+  = \sum_{\Delta\ne0} \f{\m w_\Delta\c\m w_\Delta}{\Delta} \,
+  \mean{X\otimes X}
   = \sum_{\Delta\ne0} \f{\m w_\Delta\c\m w_\Delta}{N\p{N-1}\Delta}
-  \p{\col{X}^2 - N \col{X^2}}.
+  \p{\col{X}^2 - N \col{X^2}},
   \label{eq:H_eff_2_1}
 \end{align}
-In the SU(2) case, this result implies that an inhomogeneous magnetic
-field ($X=s_\z$) yields the OAT Hamiltonian $H_{\t{OAT}}$ in
-\eqref{eq:OAT} at second order perturbation theory.
-
-\newpage
+where $\mean{X\otimes X}$ is the average of
+$\p{X\otimes X}_{ij}=X_iX_j$ over all $i\ne j$, similarly to the
+definition of $\mean{w}$ in Eq.~\eqref{eq:mean_col}.  In the SU(2)
+case, this result implies that an inhomogeneous magnetic field
+($X=s_\z$) yields the OAT Hamiltonian $H_{\t{OAT}}$ in \eqref{eq:OAT}
+at second order perturbation theory.
 
@@
 
-\red{[note: this section is not yet finished or simplified]}
+In order to find the effective Hamiltonian $H_{\t{eff}}^{(2)}$ induced
+by the $M$-body operator $\O\p{\m w,X}$ on the PS manifold $\M_0$ at
+second order in perturbation theory, we constructed states with
+definite excitation energy $\Delta>0$ with respect to $H_0$.  Rather
+than treating these excited states as ``intermediate'' or ``virtual''
+in the derivation of an effective ground-state Hamiltonian, we can
+perform numerical simulations in the joint Hilbert space of both
+ground and excited states.  When the perturbation $\O\p{\m w,X}$ is
+small, namely with operator norm
+$\norm{\O\p{\m w,X}}<\Delta_{\t{gap}}/2$, performing such simulations
+with an appropriate choice of excited states is equivalent to
+high-order ground-state perturbation theory.  The corresponding order
+of perturbation theory is determined by the choice of excited states
+in a manner that will be clarified below.  Moreover, the first-class
+treatment of excited states enables simulating the dynamics induced by
+{\it non-perturbative} additions $\O\p{\m w,X}$ to the base
+Hamiltonian $H_0$ in finite systems.
+
+In order to simulate dynamics within excited-state manifolds, we first
+need to classify the set of excited states in order to construct a
+suitable basis.  In the case of SU(2), a good starting point is to
+identify states with definite (non-negative) spin length
+$S=\set{N/2,N/2-1,\cdots}$, and spin projection
+$m_\z\in\set{S,S-1,\cdots,-S}$ onto a quantization axis.  To
+generalize this classification to SU($n$), we denote the manifold of
+states with definite fixed spin length $S$ by $\M_{N/2-S}$.  The PS
+manifold of Dicke states with $S=N/2$ is then $\M_0$, the manifold of
+single-magnon spin-wave states with $S=N/2-1$ is $\M_1$, and more
+generally the manifold of $r$-magnon states with $S=N/2-r$ is $\M_r$.
+Similarly, rather than the spin projection $m_\z$ we identify states
+by labels $m=\p{m_\up,m_\dn}$ in which $m_\mu$ denotes the occupation
+number of the single-spin state $\mu$, with $m_\up-m_\dn=m_\z$ and
+$m_\up+m_\dn=N$.
+
+The label $m$ suffices to identify the PS (Dicke) manifold
+$\M_0=\oper{span}\set{\ket{m}}$, but we need additional labels to
+uniquely identify states in the $r$-magnon manifold $\M_r$ with $r>0$.
+Single-magnon ($r=1$) spin-wave states of SU(2) spin systems can be
+written in the form $\ket{m,\m v}\equiv\O\p{\m v,\sigma_\z}\ket{m}$,
+where $\m v=\p{v_1,v_2,\cdots,v_N}$ and $\sigma_\z$ is a Pauli-$z$
+operator.  If an SU(2) spin system is translationally invariant, then
+its spin-wave eigenstates have $v_j=e^{\i k_{\m v}\c j}/\sqrt{N}$ with
+wavenumber $k_{\m v}\ne0$.  Similarly, $r$-magnon states for SU(2)
+take the form
+$\ket{m,\m v}\equiv\O\p{\m v,\sigma_\z^{\otimes r}}\ket{m}$, where now
+$\m v$ is an $r$-index tensor that encodes spatial structure and
+correlations in the excited state $\ket{m,\m v}$ (see Figure
+\ref{fig:pair_structure}).
+
+\begin{figure}
+  \centering
+  \includegraphics{pair_structure/pair_L16_A3_s1.pdf}
+  \includegraphics{pair_structure/pair_L16_A3_s2.pdf}
+  \includegraphics{pair_structure/pair_L16_A3_s3.pdf} \\[1em]
+  \includegraphics{pair_structure/pair_L16_A3_s4.pdf}
+  \includegraphics{pair_structure/pair_L16_A3_s5.pdf}
+  \includegraphics{pair_structure/pair_L16_A3_s6.pdf}
+  \includegraphics{pair_structure/pair_L16_A3_s7.pdf}
+  \caption{Matrices $\m v_\Delta=\sum_{i,j}v^{(\Delta)}_{ij}\op{i}{j}$
+    that generate two-magnon states
+    $\ket{\m v_\Delta,m} \equiv \O\p{\m v_\Delta,
+      \sigma_\z\otimes\sigma_\z} \ket{m}$ with definite excitation
+    energy $\Delta$ in a periodic 1D chain of $N=16$ spins with
+    power-law SU(2) couplings $h_{ij}=-1/\oper{dist}\p{i,j}^3$, where
+    $\oper{dist}\p{i,j}\equiv\min\p{\abs{i-j},N-\abs{i-j}}$.  Darker
+    colors correspond to larger values of $\abs*{v^{(\Delta)}_{ij}}$,
+    and a blue (red) color corresponds to a positive (negative) sign
+    of $v^{(\Delta)}_{ij}$.  In the case of SU($n$) with $n>2$, the
+    ``two-magnon'' excitation energies $\Delta$ and matrices
+    $\m v_\Delta$ remain identical, but now correspond to a degenerate
+    set of states that generated by different choice of two-body
+    operators in place of $\sigma_\z\otimes\sigma_\z$.}
+  \label{fig:pair_structure}
+\end{figure}
+
+\red{[note: this section is not yet finished]}
 
@@
 
-In order to find the effective Hamiltonian $H_{\t{eff}}^{(2)}$ induced
-by the multi-body operator $\hat\O$ on the PS manifold $\M_0$ at
-second order in perturbation theory, we constructed states with
-definite excitation energy $\Delta$ with respect to $H_0$.  These
-states are generated by the action of an $M$-body operator
-$X_{\m w_\Delta}$ on a PS state $\ket\psi\in\M_0$.  We denote the
-manifold of all states that can be generated by the action of some
-$M$-body operator, but not any $r$-body operator with $r<M$, by
-$\M_M$; we will refer to $\M_M$ as the $M$-body excitation manifold.
-We then denote the a direct sum of all $\M_r$ for $r\le M$ by
-$\col\M_M$, i.e.~$\col\M_M\equiv\bigoplus_{r\le M}\M_r$.  The manifold
-$\col\M_M$ consists of all states accessible by the action of $r$-body
-operators with $r\le M$ on PS states in $\M_0$.  These definitions are
-sketched out in Figure \ref{fig:manifold_sketch}.
+\newpage
+
+\red{[note: text below is old and unsimplified]}
+
+We denote the manifold of all states that can be generated by the
+action of some $M$-body operator, but not any $r$-body operator with
+$r<M$, by $\M_M$; we will refer to $\M_M$ as the $M$-body excitation
+manifold.  We then denote the a direct sum of all $\M_r$ for $r\le M$
+by $\col\M_M$, i.e.~$\col\M_M\equiv\bigoplus_{r\le M}\M_r$.  The
+manifold $\col\M_M$ consists of all states accessible by the action of
+$r$-body operators with $r\le M$ on PS states in $\M_0$.  These
+definitions are sketched out in Figure \ref{fig:manifold_sketch}.
 
@@
   \D_N\p{M} \equiv
-  \set{ k \in \ZZ_N^M: \t{all entries $k_i$ of $k$ are distinct } },
+  \set{ k \in \ZZ_N^M: \t{all entries $k_i$ of $k$ are distinct} },
 \end{align}
@@
 \begin{align}
-  \p{\m h - \diag\v h} \c \m w = \Delta \m w,
+  \p{\m h - \oper{diag}\v h} \c \m w = \Delta \m w,
   \label{eq:single_body_eig}
@@
 $\v h \equiv \sum_{i,j} h_{ij} \ket{i} = \sum_i h_i \ket{i}$ is the
-sum of all columns of $\m h$; and $\diag\v h$ is a matrix with $\v h$
+sum of all columns of $\m h$; and $\oper{diag}\v h$ is a matrix with $\v h$
 on the diagonal and zeroes everywhere else.  When the interaction
@@
 analytically.  In this case, the couplings $h_{ij}$ depend only on the
-separation $\abs{i-j}$, so eigenvectors of $\m h$ are therefore plane
-waves of the form
+separation $\abs{i-j}$, so eigenvectors of $\m h$ are plane waves of
+the form
 \begin{align}
@@
 $h_{0,d}=h_{0,-d}$.  The remainder of \eqref{eq:single_body_eig} that
-we need to sort out is $\diag\v h$, where all
+we need to sort out is $\oper{diag}\v h$, where all
 $h_i = \sum_{i,j}h_{ij} = \sum_d h_{0,d}$ are equal, which implies
-that $\diag\v h = \sum_d h_{0,d}$ is a scalar.  We thus find that
-\begin{align}
-  \p{\m h - \diag\v h} \c \m w_k = \Delta_k \m w_k,
+that $\oper{diag}\v h = \sum_d h_{0,d}$ is a scalar.  We thus find that
+\begin{align}
+  \p{\m h - \oper{diag}\v h} \c \m w_k = \Delta_k \m w_k,
   &&
@@
   \sp{\sum_{i\in\ZZ_N} h_{ik} \bra{i} - h_k \bra{k}}
-  = \m h - \diag\v h,
+  = \m h - \oper{diag}\v h,
 \end{align}
@@
 \end{document}
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: t
+%%% End:

### 2020-06-09 collective_sun/scripts/plot_pair_structure.py

@@ -1,0 +1,60 @@
+#!/usr/bin/env python3
+
+import numpy as np
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+
+from multibody_methods import dist_method, get_multibody_states
+
+spin_num = 16
+alpha = 3
+manifold = 2
+periodic = True
+
+font_size = 8
+figsize = (1.5,1.5)
+color_map = "RdBu"
+
+fig_dir = "../figures/pair_structure/"
+
+params = { "font.family" : "serif",
+           "font.serif" : "Computer Modern",
+           "text.usetex" : True,
+           "font.size" : font_size }
+plt.rcParams.update(params)
+
+plot_args = dict( cmap = color_map, aspect = "equal", interpolation = "nearest" )
+
+##########################################################################################
+# build SU(n) interaction matrix, and build generators of interaction eigenstates
+##########################################################################################
+
+lattice_shape = (spin_num,)
+dist = dist_method(lattice_shape, periodic = periodic)
+
+sunc = {}
+spin_num = np.prod(lattice_shape)
+sunc["mat"] = np.zeros((spin_num,spin_num))
+for pp in range(spin_num):
+    for qq in range(pp+1,spin_num):
+        sunc["mat"][pp,qq] = sunc["mat"][qq,pp] = -1/dist(pp,qq)**alpha
+sunc["TI"] = periodic
+
+sunc["shells"], sunc["energies"], sunc_tensors \
+    = get_multibody_states(lattice_shape, sunc["mat"], manifold, sunc["TI"])
+sunc.update(sunc_tensors)
+
+shells = sunc["shells"][manifold]
+max_mag = max( max(abs(sunc[shell].flatten())) for shell in shells )
+norm = mpl.colors.Normalize(vmin = -max_mag, vmax = max_mag)
+
+for shell in shells:
+    plt.figure(figsize = figsize)
+    energy = sunc["energies"][shell]
+    plt.title(r"$\Delta\approx" + f"{energy:.2f}" + "$")
+    plt.imshow(sunc[shell], norm = norm, **plot_args)
+    plt.xticks([])
+    plt.yticks([])
+    plt.tight_layout(pad = 0.1)
+    plt.savefig(fig_dir + f"pair_L{spin_num}_a{alpha}_s{shell}.pdf")
+    plt.close()

### 2020-06-10 collective_sun/sun_perturbation.tex
added option to get only several low-lying states within each r-body excitation manifold added more text describing excitation manifolds

@@
 \usepackage{tikz}
-\usetikzlibrary{arrows,math}
+\usetikzlibrary{arrows,math,calc}
 \tikzstyle{every picture}+=[remember picture]
+\tikzset{
+    thickest/.style={line width=3pt}
+}
 
@@
 small, namely with operator norm
-$\norm{\O\p{\m w,X}}<\Delta_{\t{gap}}/2$, performing such simulations
-with an appropriate choice of excited states is equivalent to
-high-order ground-state perturbation theory.  The corresponding order
-of perturbation theory is determined by the choice of excited states
-in a manner that will be clarified below.  Moreover, the first-class
-treatment of excited states enables simulating the dynamics induced by
+$\norm{\O\p{\m w,X}}<\Delta_{\t{gap}}/2$, such simulations are
+equivalent (up to perturbative corrections) to a higher-order
+perturbation theory.  Moreover, the first-class treatment of excited
+states in finite systems enables simulating the dynamics induced by
 {\it non-perturbative} additions $\O\p{\m w,X}$ to the base
-Hamiltonian $H_0$ in finite systems.
+Hamiltonian $H_0$.
 
@@
 need to classify the set of excited states in order to construct a
-suitable basis.  In the case of SU(2), a good starting point is to
-identify states with definite (non-negative) spin length
-$S=\set{N/2,N/2-1,\cdots}$, and spin projection
+suitable basis.  In the case of SU(2), a common starting point is to
+identify states with a definite (non-negative) spin length
+$S=\set{N/2,N/2-1,\cdots}$, and a spin projection
 $m_\z\in\set{S,S-1,\cdots,-S}$ onto a quantization axis.  To
@@
 generally the manifold of $r$-magnon states with $S=N/2-r$ is $\M_r$.
-Similarly, rather than the spin projection $m_\z$ we identify states
-by labels $m=\p{m_\up,m_\dn}$ in which $m_\mu$ denotes the occupation
-number of the single-spin state $\mu$, with $m_\up-m_\dn=m_\z$ and
-$m_\up+m_\dn=N$.
+Rather than the spin projection $m_\z$ we then identify states by
+labels $m=\p{m_\up,m_\dn}\in\A_2\p{N}$ in which $m_\mu$ denotes the
+occupation number of the single-spin state $\mu$, with
+$m_\up-m_\dn=m_\z$ and $m_\up+m_\dn=N$.
 
 The label $m$ suffices to identify the PS (Dicke) manifold
-$\M_0=\oper{span}\set{\ket{m}}$, but we need additional labels to
-uniquely identify states in the $r$-magnon manifold $\M_r$ with $r>0$.
-Single-magnon ($r=1$) spin-wave states of SU(2) spin systems can be
-written in the form $\ket{m,\m v}\equiv\O\p{\m v,\sigma_\z}\ket{m}$,
-where $\m v=\p{v_1,v_2,\cdots,v_N}$ and $\sigma_\z$ is a Pauli-$z$
-operator.  If an SU(2) spin system is translationally invariant, then
-its spin-wave eigenstates have $v_j=e^{\i k_{\m v}\c j}/\sqrt{N}$ with
+$\M_0=\oper{span}\set{\ket{m}:m\in\A_2\p{N}}$, but we need additional
+labels to uniquely identify states in the $r$-magnon manifold $\M_r$
+with $r>0$.  Single-magnon ($r=1$) spin-wave states of SU(2) spin
+systems can be written in the form
+\begin{align}
+  \ket{m,\m v} \equiv \O\p{\m v,\sigma_\z} \ket{m}
+  = \sum_j v_j \sigma_\z^{(j)} \ket{m},
+\end{align}
+where each $v_j$ is a scalar and $\sigma_\z$ is a Pauli-$z$ operator.
+If an SU(2) spin system is translationally invariant, then these
+spin-wave eigenstates have $v_j=e^{\i k_{\m v}\c j}/\sqrt{N}$ with
 wavenumber $k_{\m v}\ne0$.  Similarly, $r$-magnon states for SU(2)
@@
 $\ket{m,\m v}\equiv\O\p{\m v,\sigma_\z^{\otimes r}}\ket{m}$, where now
-$\m v$ is an $r$-index tensor that encodes spatial structure and
+$\m v$ is an $r$-index tensor that encodes the structure of
 correlations in the excited state $\ket{m,\m v}$ (see Figure
-\ref{fig:pair_structure}).
+\ref{fig:structure}).  Note that $M$-body operators will generally
+couple the PS manifold $\M_0$ to all $r$-magnon manifolds $\M_r$ with
+$r\le M$ (see Figure \ref{fig:manifold_sketch}).  To generalize the
+notion of $r$-magnon manifolds $\M_r$ to the case of SU($n$), where
+there is no obvious notion of a spin length $S$, we therefore define
+$\M_r$ for $r>0$ as the manifold of states that (i) can be generated
+by the action of $r$-body operators on $\M_0$, and (ii) are orthogonal
+to all states in $\bigoplus_{s<r}\M_s$.
 
@@
   \centering
-  \includegraphics{pair_structure/pair_L16_A3_s1.pdf}
-  \includegraphics{pair_structure/pair_L16_A3_s2.pdf}
-  \includegraphics{pair_structure/pair_L16_A3_s3.pdf} \\[1em]
-  \includegraphics{pair_structure/pair_L16_A3_s4.pdf}
-  \includegraphics{pair_structure/pair_L16_A3_s5.pdf}
-  \includegraphics{pair_structure/pair_L16_A3_s6.pdf}
-  \includegraphics{pair_structure/pair_L16_A3_s7.pdf}
-  \caption{Matrices $\m v_\Delta=\sum_{i,j}v^{(\Delta)}_{ij}\op{i}{j}$
-    that generate two-magnon states
-    $\ket{\m v_\Delta,m} \equiv \O\p{\m v_\Delta,
-      \sigma_\z\otimes\sigma_\z} \ket{m}$ with definite excitation
-    energy $\Delta$ in a periodic 1D chain of $N=16$ spins with
-    power-law SU(2) couplings $h_{ij}=-1/\oper{dist}\p{i,j}^3$, where
-    $\oper{dist}\p{i,j}\equiv\min\p{\abs{i-j},N-\abs{i-j}}$.  Darker
-    colors correspond to larger values of $\abs*{v^{(\Delta)}_{ij}}$,
-    and a blue (red) color corresponds to a positive (negative) sign
-    of $v^{(\Delta)}_{ij}$.  In the case of SU($n$) with $n>2$, the
-    ``two-magnon'' excitation energies $\Delta$ and matrices
-    $\m v_\Delta$ remain identical, but now correspond to a degenerate
-    set of states that generated by different choice of two-body
-    operators in place of $\sigma_\z\otimes\sigma_\z$.}
-  \label{fig:pair_structure}
+  \includegraphics{structure/mat_L101_a2.pdf}
+  \caption{Visualizations of dimension-$r$ tensors $\m v$ that
+    generate the lowest-lying $r$-magnon states
+    $\ket{\m v,m} \equiv \O\p{\m v, \sigma_\z^{\otimes{r}}} \ket{m}$
+    with definite excitation energy $\Delta$ above the PS manifold
+    $\M_0$ of a periodic 1D lattice of $N=101$ spins interacting via
+    the SU(2)-symmetric Hamiltonian in \eqref{eq:H_SU2} with power-law
+    couplings $h_{ij}=-1/\oper{dist}\p{i,j}^2$.  Here
+    $\oper{dist}\p{i,j}\equiv\min\p{\abs{i-j},N-\abs{i-j}}$ is the
+    minimal distance between spins $i,j$ on a periodic 1D lattice.
+    Top row ($r=2$) shows the full 2D matrix
+    $\m v = \sum_{i,j}v_{ij}\op{i}{j}$, and bottom row ($r=3$) shows a
+    single 2D slice of the full 3D tensor,
+    $\tilde{\m v}_0 = \sum_{i,j}v_{ij,0}\op{i}{j}$.  Blue (red) pixels
+    correspond to matrix elements with a positive (negative) sign, and
+    darker colors indicate larger absolute values.  In the case of
+    $r=2$, translational invariance enforces $v_{ij}=v_{i-d,j-d}$ for
+    all displacements $d$, so every row of $\m v$ is some translation
+    of a single (arbitrary) row.  Similarly, when $r=3$ every 2D slice
+    of the full 3D tensor can be obtained by appropriately translating
+    the 2D slice $\tilde{\m v}_0$ shown above.  In the case of SU($n$)
+    with $n>2$, the ``$r$-magnon'' excitation energies $\Delta$ and
+    tensors $\m v$ remain identical, but now correspond to a
+    degenerate set of states that are generated by different choices
+    of $r$-body operators in place of $\sigma_\z^{\otimes r}$.}
+  \label{fig:structure}
 \end{figure}
-
-\red{[note: this section is not yet finished]}
-
-% todo: give examples of these manifolds in SU(2)
-% todo: clarify connection between multi-body excitation manifolds
-%       and manifolds of fixed total spin in the SU(2) case
-
-\newpage
-
-\red{[note: text below is old and unsimplified]}
-
-We denote the manifold of all states that can be generated by the
-action of some $M$-body operator, but not any $r$-body operator with
-$r<M$, by $\M_M$; we will refer to $\M_M$ as the $M$-body excitation
-manifold.  We then denote the a direct sum of all $\M_r$ for $r\le M$
-by $\col\M_M$, i.e.~$\col\M_M\equiv\bigoplus_{r\le M}\M_r$.  The
-manifold $\col\M_M$ consists of all states accessible by the action of
-$r$-body operators with $r\le M$ on PS states in $\M_0$.  These
-definitions are sketched out in Figure \ref{fig:manifold_sketch}.
 
@@
   \begin{tikzpicture}
-    \draw[ultra thick] (0,0) -- node[above] {$\M_0$} ++(10em,0);
-    \draw[ultra thick] (0,3em) -- node[above] {$\M_1$} ++(10em,0);
-    \draw[ultra thick] (0,2*3em) -- node[above] {$\M_2$} ++(10em,0);
-    \draw[->] (-.2em,0) to [bend left] (-.2em,3em);
-    \node (X) at (.2em,0.5*3em) {$\hat\O_1$};
-    \draw[->] (-.4em,0) to [bend left] (-.2em,2*3em);
-    \node (XY) at (-2em,3em) {$\hat\O_2$};
-    \draw (10em+.2em,-.2em) -- (10em+.4em,-.2em)
-    -- (10em+.4em,3em+.2em) -- (10em+.2em,3em+.2em);
-    \node (cM1) at (10em+1.5em,0.5*3em) {$\col\M_1$};
-    \draw (10em+2.4em+.2em,-.4em) -- (10em+2.4em+.4em,-.4em)
-    -- (10em+2.4em+.4em,6em+.4em) -- (10em+2.4em+.2em,6em+.4em);
-    \node (cM1) at (10em+2.4em+1.5em,3em) {$\col\M_2$};
+    \draw[thickest] (0,0) -- ++(10em,0);
+    \foreach \hh in {0.15em,0.05em,-0.05em,-0.15em} {
+      \draw (0,3em+\hh) -- ++(10em,0);
+    }
+    \draw[->] (7em,.25em) arc (220:0:.4em);
+    \draw[->] (7em,.25em) to [bend left]
+    node[left] {$\O\p{\m v,\sigma_\z}$} (7em,2.7em);
+    \node () at (5em,4.7em) {$\vdots$};
+    \foreach \hh in {0.25em,0.15em,0.05em,-0.05em,-0.15em,-0.25em} {
+      \draw (0,6em+\hh) -- ++(10em,0);
+    }
+    \draw[->] (-.4em,0em) arc (220:0:.4em);
+    \draw[->] (-.4em,0) to [bend left, in=130] (-.2em,3em);
+    \draw[->] (-.4em,0) to [bend left, in=130] (-.2em,4.3em);
+    \draw[->] (-.4em,0) to [bend left, in=130]
+    node[left] {$\O\p{\m w,\sigma_\z^{\otimes M}}$} (-.2em,6em);
+    \node[right] () at (10em,0) {$\M_0,S=\frac{N}{2}$};
+    \node[right] () at (10em,3em) {$\M_1,S=\frac{N}{2}-1$};
+    \node[right] () at (10em,6em) {$\M_M,S=\frac{N}{2}-M$};
   \end{tikzpicture}
-  \caption{Schematic of low-lying multi-body excitation manifolds.
-    Starting with the permutationally-state manifold $\M_0$, the
-    single-body excitation manifold $\M_1$ consists of states
-    accessible via single-body perturbations $\hat\O_1$.  The two-body
-    excitation manifold $\hat\O_2$ then consists of states accessible
-    via two-body perturbations, but not single-body perturbations.
-    The joint manifold $\col\M_M$ consists of all states accessible by
-    operators $r$-body operators $\O_r$ with $r\le M$.}
+  \caption{Schematic of relationships between $r$-magnon manifolds
+    $\M_r$ with total spin $S=N/2-r$ in an SU(2) spin system.
+    Single-body operators $\O\p{\m v,\sigma_\z}$ generally couple the
+    ground-state manifold $\M_0$ to the single-magnon manifold $\M_1$
+    as well as $\M_0$ itself, while $M$-body operators
+    $\O\p{\m w,\sigma_\z^{\otimes M}}$ couple $\M_0$ to all $r$-magnon
+    manifolds $\M_r$ with $r\le M$.}
   \label{fig:manifold_sketch}
 \end{figure}
+
+\red{[note: this section is not yet finished]}
+
+\newpage
+
+\red{[note: the remainder of this section is old, and will be replaced
+  entirely]}
 
@@
 
-Here we show that power-law SU($n$)-symmetric interactions of the form
-in \eqref{eq:H_0} with $h_{ij}=-h/\abs{i-j}^\alpha$ yield a
-non-vanishing many-body interaction energy gap when $\alpha\le D$,
-where $D$ is the dimension of the lattice.  On a periodic lattice of
-$N=L^D$ spins, the spectral gap of the interaction Hamiltonian is (see
-Appendix \ref{sec:single_body_eigenstates})
+Here we show that in $D$ spatial dimensions, power-law
+SU($n$)-symmetric interactions of the form in \eqref{eq:H_0} with
+$h_{ij}=-1/\abs{i-j}^\alpha$ yield a non-vanishing spectral gap when
+$\alpha\le D$.  For definiteness, we will consider a periodic
+hypercubic lattice of $N=L^D$ spins.  The spectral gap of the
+interaction Hamiltonian is then (see Appendix
+\ref{sec:single_body_eigenstates})
 \begin{align}
@@
   = \sum_{d\in\ZZ_L^D} h_{0,d} \sp{\cos\p{d \c k_{\t{SE}}}-1}
-  = h \sum_{\substack{d\in\ZZ_L^D\\\abs{d}\ge1}}
+  = \sum_{\substack{d\in\ZZ_L^D\\\abs{d}\ge1}}
   \f{1-\cos\p{d \c k_{\t{SE}}}}{\abs{d}^\alpha},
 \end{align}
-where $\ZZ_L$ denotes the integers modulo $L$;
-$k_{\t{SE}}\in\ZZ_L^D\times2\pi/L$ is a wavenumber for a
-singly-excited spin-wave state; and the norm $\abs{d}$ for a vector
-$d$ on a periodic lattice is implicitly understood to mean the
-smallest Euclidean distance of $d$ from a fixed origin.  In order to
-yield the smallest excitation energy $\Delta$, the wavenumber
-$k_{\t{SE}}$ should maximize the contribution of the cosine term
-above, which is achieved by a wavenumber that minimizes the
-oscillations of this term when integrated over the entire lattice.  A
-suitable candidate for a minimal wavenumber is
-$k_{\t{SE}}=\p{2\pi/L,0,0,\cdots}$, which corresponds to an excitation
-energy
-\begin{align}
-  \Delta = h \sum_{\substack{d\in\ZZ_L^D\\\abs{d}\ge1}}
+where $\ZZ_L$ denotes the set of integers modulo $L$;
+$k_{\t{SE}}\in\ZZ_L^D\times2\pi/L$ is a wavenumber for a single-magnon
+spin-wave state; and the norm $\abs{d}$ for a vector $d$ on a periodic
+lattice is implicitly understood to mean the smallest Euclidean
+distance of $d$ from a fixed origin.  In order to yield the smallest
+excitation energy $\Delta$, the wavenumber $k_{\t{SE}}$ should
+maximize the contribution of the cosine term above, which is achieved
+by a wavenumber that minimizes the oscillations of this term when
+integrated over the entire lattice.  A suitable candidate for a
+minimal wavenumber is $k_{\t{SE}}=\p{2\pi/L,0,0,\cdots}$, which
+corresponds to an excitation energy
+\begin{align}
+  \Delta = \sum_{\substack{d\in\ZZ_L^D\\\abs{d}\ge1}}
   \f{1-\cos\p{d_12\pi/L}}{\abs{d}^\alpha}.
@@
 get
-% todo: define \simeq ?
 \begin{align}
   \Delta
-  = h \sum_{\substack{x\in\SS_\epsilon^D\\\abs{x}\ge\epsilon}}
+  = \sum_{\substack{x\in\SS_\epsilon^D\\\abs{x}\ge\epsilon}}
   \f{1-\cos\p{\pi x_1}}{\abs{x/\epsilon}^\alpha}
-  = h \epsilon^{\alpha-D} \sum_{\substack{x\in\SS_L^D\\\abs{x}\ge\epsilon}}
+  = \epsilon^{\alpha-D} \sum_{\substack{x\in\SS_L^D\\\abs{x}\ge\epsilon}}
   \epsilon^D \f{1-\cos\p{\pi x_1}}{\abs{x}^\alpha}.
@@
 \begin{align}
-  \Delta = h \epsilon^{\alpha-D} \I_D\p{\epsilon},
+  \Delta = \epsilon^{\alpha-D} \I_D\p{\epsilon},
   &&
@@
 \end{align}
-where the integral $\I_D\p{\epsilon}$ is defined using the interval
-$\TT_a\equiv\p{-a,a}$.  The integrand of $\I_D\p{\epsilon}$ is
-strictly positive and well-behaved on the entirety of its domain
-except for the origin, where depending on the value of $\alpha$ the
-integrand may vanish or diverge as $\abs{x}\to0$.  Together, these
-facts mean that
+where the integral $\I_D\p{\epsilon}$ is defined using the symmetric
+interval $\TT_a\equiv\p{-a/2,a/2}$.  The integrand of
+$\I_D\p{\epsilon}$ is strictly positive and well-behaved on the
+entirety of its domain except for the origin, where depending on the
+value of $\alpha$ the integrand may vanish or diverge as
+$\abs{x}\to0$.  Together, these facts mean that
 \begin{align}
@@
   &&
-  \Delta \stackrel{\epsilon\to0}{\sim} h \epsilon^{\alpha-D-\gamma},
+  \Delta \stackrel{\epsilon\to0}{\sim} \epsilon^{\alpha-D-\gamma},
 \end{align}
@@
 \end{align}
-where
+where for convenience we define
 \begin{align}
@@
 \end{align}
-is the set of all subsets (``choices'') of $q$ elements from $\ZZ_N$.
+as the set of all subsets (``choices'') of $q$ elements from $\ZZ_N$.
 The sum in \eqref{eq:diagnosis_start} has terms with $i,j\notin k$,
@@
   will suppress the explicit dependence of $\S_M$ on $N$, as well as
-  the dependence of $\S_M$ on any other factors such as lattice
-  geometry.}, where elements $s\in\S_M$ are subsets
-$s\subset\D_N\p{M}$ that are invariant under the appropriate symmetry
-group action.  In the case of $M=2$ in a translationally invariant and
-isotropic system, for example, each equivalence class $s\in\S_2$ would
-consist of all pairs $\p{i,j}\in\D_N\p{2}$ that are a fixed distance
-$\abs{i-j}=d_s$ apart,
+  other factors such as lattice geometry.}, where elements $s\in\S_M$
+are subsets $s\subset\D_N\p{M}$ that are invariant under the
+appropriate symmetry group action.  In the case of $M=2$ in a
+translationally invariant and isotropic system, for example, each
+equivalence class $s\in\S_2$ would consist of all pairs
+$\p{i,j}\in\D_N\p{2}$ that are a fixed distance $\abs{i-j}=d_s$ apart,
 i.e.~$s=\set{\p{i,j}\in\D_N\p{2}:\abs{i-j}=d_s}$.  If the tensor
@@
 \end{align}
-where all $w_k$ with $k\in s$ are equal by assumption, so we can
-define $w_s\equiv w_k$ for any $k\in s$.  Finally, we expand the
-projection of $\hat{\m h}$ onto the symmetric subspace:
+where all $w_k$ with $k\in s$ are equal by assumption, so
+$w_s\equiv\sum_{k\in s}w_k/\abs{s}=w_\ell$ for any $\ell\in s$.
+Finally, we expand the projection of $\hat{\m h}$ onto the symmetric
+subspace:
 \begin{align}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The general case: a diagrammatic expansion}
+\subsection{A diagrammatic expansion}
 

### 2020-06-10 collective_sun/scripts/multibody_methods.py

@@ -318,7 +318,7 @@ def multibody_problem(lattice_shape, sun_coefs, dimension, TI = None, isotropic 
 #     (2) energies of each shell, and
 #     (4) tensors that generate states in that shell
 def get_multibody_states(lattice_shape, sun_coefs, manifolds, TI, isotropic = None,
-                         updates = True):
+                         updates = True, shell_cutoff = None):
     site_num = sun_coefs.shape[0]
 
     if type(manifolds) is int:
@@ -362,6 +362,8 @@ def get_multibody_states(lattice_shape, sun_coefs, manifolds, TI, isotropic = No
             if updates:
                 print("  shells:", shell_num)
                 sys.stdout.flush()
+            if shell_cutoff and shell_num - old_shell_num == shell_cutoff:
+                break
 
         manifold_shells[manifold] = np.array(range(old_shell_num,shell_num), dtype = int)
 

### 2020-06-10 collective_sun/scripts/plot_pair_structure.py

@@ -1,60 +1,0 @@
-#!/usr/bin/env python3
-
-import numpy as np
-import matplotlib as mpl
-import matplotlib.pyplot as plt
-
-from multibody_methods import dist_method, get_multibody_states
-
-spin_num = 16
-alpha = 3
-manifold = 2
-periodic = True
-
-font_size = 8
-figsize = (1.5,1.5)
-color_map = "RdBu"
-
-fig_dir = "../figures/pair_structure/"
-
-params = { "font.family" : "serif",
-           "font.serif" : "Computer Modern",
-           "text.usetex" : True,
-           "font.size" : font_size }
-plt.rcParams.update(params)
-
-plot_args = dict( cmap = color_map, aspect = "equal", interpolation = "nearest" )
-
-##########################################################################################
-# build SU(n) interaction matrix, and build generators of interaction eigenstates
-##########################################################################################
-
-lattice_shape = (spin_num,)
-dist = dist_method(lattice_shape, periodic = periodic)
-
-sunc = {}
-spin_num = np.prod(lattice_shape)
-sunc["mat"] = np.zeros((spin_num,spin_num))
-for pp in range(spin_num):
-    for qq in range(pp+1,spin_num):
-        sunc["mat"][pp,qq] = sunc["mat"][qq,pp] = -1/dist(pp,qq)**alpha
-sunc["TI"] = periodic
-
-sunc["shells"], sunc["energies"], sunc_tensors \
-    = get_multibody_states(lattice_shape, sunc["mat"], manifold, sunc["TI"])
-sunc.update(sunc_tensors)
-
-shells = sunc["shells"][manifold]
-max_mag = max( max(abs(sunc[shell].flatten())) for shell in shells )
-norm = mpl.colors.Normalize(vmin = -max_mag, vmax = max_mag)
-
-for shell in shells:
-    plt.figure(figsize = figsize)
-    energy = sunc["energies"][shell]
-    plt.title(r"$\Delta\approx" + f"{energy:.2f}" + "$")
-    plt.imshow(sunc[shell], norm = norm, **plot_args)
-    plt.xticks([])
-    plt.yticks([])
-    plt.tight_layout(pad = 0.1)
-    plt.savefig(fig_dir + f"pair_L{spin_num}_a{alpha}_s{shell}.pdf")
-    plt.close()

### 2020-06-10 collective_sun/scripts/plot_structure.py

@@ -1,0 +1,105 @@
+#!/usr/bin/env python3
+
+import numpy as np
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+
+from multibody_methods import dist_method, get_multibody_states
+
+spin_num = 101
+alpha = 2
+columns = 4
+shell_cutoff = 20
+
+font_size = 9
+color_map = "RdBu"
+
+panel_figsize = (6,3.3)
+single_figsize = (1.5,1.5)
+
+fig_dir = "../figures/structure/"
+
+params = { "font.family" : "serif",
+           "font.serif" : "Computer Modern",
+           "text.usetex" : True,
+           "font.size" : font_size,
+           "axes.titlesize" : font_size }
+plt.rcParams.update(params)
+
+norm = mpl.colors.Normalize(vmin = -1, vmax = 1)
+plot_args = dict( cmap = color_map, norm = norm,
+                  origin = "lower", interpolation = "none", aspect = "equal" )
+
+##########################################################################################
+# build SU(n) interaction matrix, and build generators of interaction eigenstates
+##########################################################################################
+
+max_manifold = 3
+periodic = True
+lattice_shape = (spin_num,)
+dist = dist_method(lattice_shape, periodic = periodic)
+
+sunc = {}
+spin_num = np.prod(lattice_shape)
+sunc["mat"] = np.zeros((spin_num,spin_num))
+for pp in range(spin_num):
+    for qq in range(pp+1,spin_num):
+        sunc["mat"][pp,qq] = sunc["mat"][qq,pp] = -1/dist(pp,qq)**alpha
+sunc["TI"] = periodic
+
+sunc["shells"], sunc["energies"], sunc_tensors \
+    = get_multibody_states(lattice_shape, sunc["mat"],
+                           max_manifold, sunc["TI"], shell_cutoff = shell_cutoff)
+sunc.update(sunc_tensors)
+
+shell_num = len(sunc["energies"])
+shell_digits = len(str(shell_num))
+
+##########################################################################################
+# plot generators within the 2- and 3-excitation manifolds
+##########################################################################################
+
+# make figure showing several shells
+figure, axes = plt.subplots(2, columns, figsize = panel_figsize)
+
+for manifold, row_axes in zip([ 2, 3 ], axes):
+
[…]

### 2020-06-14 collective_sun/sun_perturbation.tex
added high-level description of the construction of excited states in the general SU(n) case

@@
 
-\title{SU($n$) ferromagnets near the Heisenberg critical point}%
-\author{Michael A. Perlin}%
+\title{SU($n$) ferromagnets near the Heisenberg point}
+\author{Michael A. Perlin}
 \date{\today}
@@
 If these interactions are sufficiently weak for perturbative
-treatment, as in the case of the XXZ model near the Heisenberg
-critical point, then the effective Hamiltonian induced by
-$H_{\t{Ising}}$ on initially PS states $\ket\psi\in\M_0$ is the
-one-axis twisting (OAT) Hamiltonian\cite{ma2011quantum}
+treatment, as in the case of the XXZ model near the Heisenberg point,
+then the effective Hamiltonian induced by $H_{\t{Ising}}$ on initially
+PS states $\ket\psi\in\M_0$ is the one-axis twisting (OAT)
+Hamiltonian\cite{ma2011quantum}
 \begin{align}
@@
 $\norm{\O\p{\m w,X}}<\Delta_{\t{gap}}/2$, such simulations are
-equivalent (up to perturbative corrections) to a higher-order
-perturbation theory.  Moreover, the first-class treatment of excited
-states in finite systems enables simulating the dynamics induced by
-{\it non-perturbative} additions $\O\p{\m w,X}$ to the base
+equivalent (up to vanishing perturbative corrections) to a
+higher-order perturbation theory.  Moreover, the first-class treatment
+of excited states in finite systems enables simulating the dynamics
+induced by {\it non-perturbative} additions $\O\p{\m w,X}$ to the base
 Hamiltonian $H_0$.
@@
 In order to simulate dynamics within excited-state manifolds, we first
-need to classify the set of excited states in order to construct a
-suitable basis.  In the case of SU(2), a common starting point is to
-identify states with a definite (non-negative) spin length
+need to classify the set of excited states, and construct a suitable
+basis.  In the case of SU(2), a common starting point is to identify
+states with a definite (non-negative) spin length
 $S=\set{N/2,N/2-1,\cdots}$, and a spin projection
@@
 occupation number of the single-spin state $\mu$, with
-$m_\up-m_\dn=m_\z$ and $m_\up+m_\dn=N$.
+$m_\up+m_\dn=N$ and $m_\up-m_\dn=m_\z$.
+
+\begin{figure}
+  \centering
+  \includegraphics{structure/mat_L101_a2.pdf}
+  \caption{Visualizations of dimension-$r$ tensors $\m w$ that
+    generate the lowest-lying (unnormalized) $r$-magnon states
+    $\ket{\m w,m} \equiv \O\p{\m w, \sigma_\z^{\otimes{r}}} \ket{m}$
+    with definite excitation energy $\Delta$ above the PS manifold
+    $\M_0$ of a periodic 1D lattice of $N=101$ spins interacting via
+    the SU(2)-symmetric Hamiltonian in \eqref{eq:H_SU2} with power-law
+    couplings $h_{ij}=-1/\oper{dist}\p{i,j}^2$.  Here
+    $\oper{dist}\p{i,j}\equiv\min\p{\abs{i-j},N-\abs{i-j}}$ is the
+    minimal distance between spins $i,j$.  Top row ($r=2$) shows the
+    full 2D matrix $\m w = \sum_{i,j}w_{ij}\op{i}{j}$, and bottom row
+    ($r=3$) shows a single 2D slice of the full 3D tensor,
+    $\tilde{\m w}_0 = \sum_{i,j}w_{ij,0}\op{i}{j}$.  Blue (red) pixels
+    correspond to matrix elements with a positive (negative) sign, and
+    darker colors indicate larger absolute values.  In the case of
+    $r=2$, translational invariance enforces $w_{ij}=w_{i-d,j-d}$ for
+    all displacements $d$, so every row of $\m w$ is some translation
+    of a single (arbitrary) row.  Similarly, when $r=3$ every 2D slice
+    of the full 3D tensor can be obtained by appropriately translating
+    the 2D slice $\tilde{\m w}_0$ shown above.  In the case of SU($n$)
+    with $n>2$, the ``$r$-magnon'' excitation energies $\Delta$ and
+    tensors $\m w$ remain identical, but now correspond to a
+    degenerate set of states that are generated by different choices
+    of $r$-body operators in place of $\sigma_\z^{\otimes r}$.}
+  \label{fig:structure}
+\end{figure}
 
@@
 If an SU(2) spin system is translationally invariant, then these
-spin-wave eigenstates have $v_j=e^{\i k_{\m v}\c j}/\sqrt{N}$ with
-wavenumber $k_{\m v}\ne0$.  Similarly, $r$-magnon states for SU(2)
-take the form
-$\ket{m,\m v}\equiv\O\p{\m v,\sigma_\z^{\otimes r}}\ket{m}$, where now
-$\m v$ is an $r$-index tensor that encodes the structure of
-correlations in the excited state $\ket{m,\m v}$ (see Figure
-\ref{fig:structure}).  Note that $M$-body operators will generally
-couple the PS manifold $\M_0$ to all $r$-magnon manifolds $\M_r$ with
-$r\le M$ (see Figure \ref{fig:manifold_sketch}).  To generalize the
-notion of $r$-magnon manifolds $\M_r$ to the case of SU($n$), where
-there is no obvious notion of a spin length $S$, we therefore define
-$\M_r$ for $r>0$ as the manifold of states that (i) can be generated
-by the action of $r$-body operators on $\M_0$, and (ii) are orthogonal
-to all states in $\bigoplus_{s<r}\M_s$.
-
-\begin{figure}
-  \centering
-  \includegraphics{structure/mat_L101_a2.pdf}
-  \caption{Visualizations of dimension-$r$ tensors $\m v$ that
-    generate the lowest-lying $r$-magnon states
-    $\ket{\m v,m} \equiv \O\p{\m v, \sigma_\z^{\otimes{r}}} \ket{m}$
-    with definite excitation energy $\Delta$ above the PS manifold
-    $\M_0$ of a periodic 1D lattice of $N=101$ spins interacting via
-    the SU(2)-symmetric Hamiltonian in \eqref{eq:H_SU2} with power-law
-    couplings $h_{ij}=-1/\oper{dist}\p{i,j}^2$.  Here
-    $\oper{dist}\p{i,j}\equiv\min\p{\abs{i-j},N-\abs{i-j}}$ is the
-    minimal distance between spins $i,j$ on a periodic 1D lattice.
-    Top row ($r=2$) shows the full 2D matrix
-    $\m v = \sum_{i,j}v_{ij}\op{i}{j}$, and bottom row ($r=3$) shows a
-    single 2D slice of the full 3D tensor,
-    $\tilde{\m v}_0 = \sum_{i,j}v_{ij,0}\op{i}{j}$.  Blue (red) pixels
-    correspond to matrix elements with a positive (negative) sign, and
-    darker colors indicate larger absolute values.  In the case of
-    $r=2$, translational invariance enforces $v_{ij}=v_{i-d,j-d}$ for
-    all displacements $d$, so every row of $\m v$ is some translation
-    of a single (arbitrary) row.  Similarly, when $r=3$ every 2D slice
-    of the full 3D tensor can be obtained by appropriately translating
-    the 2D slice $\tilde{\m v}_0$ shown above.  In the case of SU($n$)
-    with $n>2$, the ``$r$-magnon'' excitation energies $\Delta$ and
-    tensors $\m v$ remain identical, but now correspond to a
-    degenerate set of states that are generated by different choices
-    of $r$-body operators in place of $\sigma_\z^{\otimes r}$.}
-  \label{fig:structure}
-\end{figure}
+spin-wave eigenstates have $v_j=e^{\i k_{\m v}\c j}$ with wavenumber
+$k_{\m v}\ne0$.  Similarly, $r$-magnon states for SU(2) take the form
+$\ket{m,\m w}\equiv\O\p{\m w,\sigma_\z^{\otimes r}}\ket{m}$, where now
+$\m w$ is an $r$-index tensor that encodes the structure of
+correlations in the excited state $\ket{m,\m w}$ (see Figure
+\ref{fig:structure}).  Considering states of this form, built using
+tensor products of Pauli-$z$ ($\sigma_\z$) operators, suffices to
+construct a basis for all excited states of an SU(2) spin system (see
+Appendix \ref{sec:operator_content}).  Note that $r$-body operators
+with arbitrary coefficients $\m w$ will generally couple the PS
+manifold $\M_0$ to all $q$-magnon manifolds $\M_q$ with $q\le r$ (see
+Figure \ref{fig:manifold_sketch}).  In order to generalize the notion
+of $r$-magnon manifolds $\M_r$ to the case of SU($n$) with $n>2$,
+where the analogous notion of a spin length $S$ is less intuitive, we
+therefore define $\M_r$ for $r>0$ as the manifold of states that (i)
+can be generated by the action of $r$-body operators on $\M_0$, and
+(ii) are orthogonal to all states in $\bigoplus_{q<r}\M_q$.
 
@@
     }
-    \draw[->] (-.4em,0em) arc (220:0:.4em);
+    \draw[->] (-.4em,0) arc (220:0:.4em);
     \draw[->] (-.4em,0) to [bend left, in=130] (-.2em,3em);
@@
     \draw[->] (-.4em,0) to [bend left, in=130]
-    node[left] {$\O\p{\m w,\sigma_\z^{\otimes M}}$} (-.2em,6em);
+    node[left] {$\O\p{\m w,\sigma_\z^{\otimes r}}$} (-.2em,6em);
     \node[right] () at (10em,0) {$\M_0,S=\frac{N}{2}$};
     \node[right] () at (10em,3em) {$\M_1,S=\frac{N}{2}-1$};
-    \node[right] () at (10em,6em) {$\M_M,S=\frac{N}{2}-M$};
+    \node[right] () at (10em,6em) {$\M_r,S=\frac{N}{2}-r$};
   \end{tikzpicture}
-  \caption{Schematic of relationships between $r$-magnon manifolds
-    $\M_r$ with total spin $S=N/2-r$ in an SU(2) spin system.
+  \caption{Schematic of relationships between $q$-magnon manifolds
+    $\M_q$ with total spin $S=N/2-q$ in an SU(2) spin system.
     Single-body operators $\O\p{\m v,\sigma_\z}$ generally couple the
     ground-state manifold $\M_0$ to the single-magnon manifold $\M_1$
-    as well as $\M_0$ itself, while $M$-body operators
-    $\O\p{\m w,\sigma_\z^{\otimes M}}$ couple $\M_0$ to all $r$-magnon
-    manifolds $\M_r$ with $r\le M$.}
+    as well as $\M_0$ itself, while $r$-body operators
+    $\O\p{\m w,\sigma_\z^{\otimes r}}$ couple $\M_0$ to all $q$-magnon
+    manifolds $\M_q$ with $q\le r$.}
   \label{fig:manifold_sketch}
@@
 
-\red{[note: this section is not yet finished]}
+In the case of an SU($n$) spin system, we can similarly start with the
+ground-state PS manifold
+$\M_0=\oper{span}\set{\ket{m}:m\in\A_n\p{N}}$, and generate excited
+states by the action of multi-body operators, $\O\p{\m w,X}\ket{m}$.
+The choice of tensors $\m w$ for generating states in the $r$-magnon
+manifold $\M_r$ is the same as before, but we now have the additional
+complication of having to choose $r$-spin operators, $X$.  As with the
+choice of $\sigma_\z^{\otimes r}$ for SU(2), it suffices to choose
+tensor products of {\it diagonal} single-spin operators (that is,
+diagonal in a fixed single-particle basis), but there will now be
+$\sim\p{n-1}^r$ distinct choices that generate degenerate, orthogonal
+states: one of $n-1$ diagonal operators (that are linearly independent
+of each other and the identity) for each of $r$ tensor factors.  We
+discuss the exact choice of operators in more detail in Appendix
+\ref{sec:operator_content}.  To summarize, for each PS state
+$\ket{m}$, we identify a basis $\set{Z_{m\nu}:\nu\in\ZZ_{n-1}}$ of
+single-spin operators that are diagonal and linearly independent of
+the identity, and choose a list of integers
+$z\equiv\p{z_1,z_2,\cdots,z_{n-1}}\in\A_{n-1}\p{r}$ to define the
+$r$-spin operator
+\begin{align}
+  X_{mz} \equiv \bigotimes_{\nu=1}^{n-1} Z_{m\nu}^{\otimes z_\nu}
+  = \underbrace{Z_{m,1} \otimes \cdots \otimes Z_{m,1}}_{z_1~\t{times}}
+  \otimes
+  \underbrace{Z_{m,2} \otimes \cdots \otimes Z_{m,2}}_{z_2~\t{times}}
+  \otimes \cdots \otimes
+  \underbrace{Z_{m,n-1} \otimes \cdots
+    \otimes Z_{m,n-1}}_{z_{n-1}~\t{times}}.
+\end{align}
+Our basis for $\M_r$ then consists of the states
+$\ket{m,\m w,z}\equiv\O\p{\m w,X_{mz}}\ket{m}$.
+
+\red{[todo: discuss the use of this basis in numerical simulations]}
 
@@
 \red{[look at \cite{rai1988correlated} for classifying asymmetric
-  states]}
+  states?]}
 
@@
 \red{[TODO: write section]}
+
+\begin{acknowledgments}
+  We thank Asier Pi\~neiro Orioli for helpful discussions and
+  technical contributions to this work.
+\end{acknowledgments}
+
+\bibliography{\jobname.bib}
 
@@
 \label{sec:spectral_gap}
-
-\red{[todo: make this a subsection of Appendix
-  \ref{sec:single_body_eigenstates}]}
 
@@
 hypercubic lattice of $N=L^D$ spins.  The spectral gap of the
-interaction Hamiltonian is then (see Appendix
+interaction Hamiltonian is then determined by well-known spin-wave
+theory (see Ref.~[\citenum{ashcroft1978solid}] or Appendix
 \ref{sec:single_body_eigenstates})
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Choosing the operator content}
+\label{sec:operator_content}
+
+\red{[todo: write section]}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \abs{\L\p{g}}$.  To keep mathematical expressions simple, we will use
-symmetric tensors in all examples in this section, but our discussions
-apply just as well to the asymmetric case.
+symmetric tensors in all examples throughout this section, but our
+discussions apply just as well to the asymmetric case.
 
@@
 
-\bibliography{\jobname.bib}
-
 \end{document}

### 2020-06-17 collective_sun/sun_perturbation.tex
added appendix about choosing the operator content for excitation energy eigenstates

@@
 \newcommand{\Q}{\mathcal{Q}}
+\newcommand{\R}{\mathcal{R}}
 \renewcommand{\S}{\mathcal{S}}
 \newcommand{\T}{\mathcal{T}}
+\newcommand{\Z}{\mathcal{Z}}
 
@@
 \item an initial state in the manifold of permutationally-symmetric
-  states, a multi-level generalization of the Dicke manifold for
+  states, a multilevel generalization of the Dicke manifold for
   two-level spins, together with the assumption that
@@
 total, then Dicke states $\ket{m_\up,m_\dn}$ must have
-$m_\up+m_\dn=N$.  In the case of multi-level spins, the PS manifold
+$m_\up+m_\dn=N$.  In the case of multilevel spins, the PS manifold
 $\M_0$ is similarly spanned by states
@@
     with definite excitation energy $\Delta$ above the PS manifold
-    $\M_0$ of a periodic 1D lattice of $N=101$ spins interacting via
+    $\M_0$ in a periodic 1D lattice of $N=101$ spins interacting via
     the SU(2)-symmetric Hamiltonian in \eqref{eq:H_SU2} with power-law
@@
 diagonal in a fixed single-particle basis), but there will now be
-$\sim\p{n-1}^r$ distinct choices that generate degenerate, orthogonal
-states: one of $n-1$ diagonal operators (that are linearly independent
-of each other and the identity) for each of $r$ tensor factors.  We
-discuss the exact choice of operators in more detail in Appendix
-\ref{sec:operator_content}.  To summarize, for each PS state
-$\ket{m}$, we identify a basis $\set{Z_{m\nu}:\nu\in\ZZ_{n-1}}$ of
-single-spin operators that are diagonal and linearly independent of
-the identity, and choose a list of integers
-$z\equiv\p{z_1,z_2,\cdots,z_{n-1}}\in\A_{n-1}\p{r}$ to define the
-$r$-spin operator
-\begin{align}
-  X_{mz} \equiv \bigotimes_{\nu=1}^{n-1} Z_{m\nu}^{\otimes z_\nu}
-  = \underbrace{Z_{m,1} \otimes \cdots \otimes Z_{m,1}}_{z_1~\t{times}}
+$\sim\p{n-1}^r$ choices: one of $n-1$ diagonal operators (that are
+linearly independent of the identity) for each of $r$ tensor factors.
+We discuss the exact choice of operators in more detail in Appendix
+\ref{sec:operator_content}.  To summarize, we define $n$ single-spin
+operators $Z_\nu\equiv1-2\op\nu=\sum_{\mu\ne\nu}\op\mu-\op\nu$, for
+which $Z_\nu=\pm\sigma_\z$ in the case of SU(2), and we choose a list
+of non-negative integers $z\equiv\p{z_1,z_2,\cdots,z_n}$ with
+$\sum_\mu z_\mu=r$ to define the $r$-spin operator
+\begin{align}
+  X_z \equiv \bigotimes_{\nu=1}^n Z_\nu^{\otimes z_\nu}
+  = \underbrace{Z_1 \otimes \cdots \otimes Z_1}_{z_1~\t{times}}
   \otimes
-  \underbrace{Z_{m,2} \otimes \cdots \otimes Z_{m,2}}_{z_2~\t{times}}
+  \underbrace{Z_2 \otimes \cdots \otimes Z_2}_{z_2~\t{times}}
   \otimes \cdots \otimes
-  \underbrace{Z_{m,n-1} \otimes \cdots
-    \otimes Z_{m,n-1}}_{z_{n-1}~\t{times}}.
-\end{align}
-Our basis for $\M_r$ then consists of the states
-$\ket{m,\m w,z}\equiv\O\p{\m w,X_{mz}}\ket{m}$.
+  \underbrace{Z_n \otimes \cdots \otimes Z_n}_{z_n~\t{times}}.
+\end{align}
+Our starting basis for $\M_r$ then consists of states
+$\ket{m,\m w,z}\equiv\O\p{\m w,X_z}\ket{m}$, where we enforce that
+$z_\nu\le m_\nu$ for all $\nu$ (where $m_\nu$ is the occupation number
+of the single-spin state $\nu$), and that $z$ is zero in an entry
+where $m$ is maximal: $z_\nu=0$ for
+$\nu=\min\set{\mu:m_\mu=\max\p{m}}$.  This second restriction is
+equivalent to removing an operator from $\set{Z_\nu:\nu\in\ZZ_n}$,
+making the remaining set linearly independent of the identity
+operator.  The basis $\set{\ket{m,\m w,z}}$ for $\M_r$ is generally
+complete, but {\it not} orthonormal, so as a final step we construct
+an orthonormal basis $\set{\ket{m,\m w,X}}$, where for operators $X$
+we define $\ket{m,\m w,X}\equiv\O\p{\m w,X}\ket{m}$ (see Appendix
+\ref{sec:operator_content}).
 
@@
 
-\red{[todo: write section]}
+Throughout our construction of excitation energy eigenstates of the
+form $\O\p{\m w,X}\ket\psi$, where $\ket\psi\in\M_0$ is a PS state, we
+have thus far primarily considered the choice of scalar coefficients,
+$\m w$.  Here, we discuss the choice of operators, $X$, which allows
+us to construct a complete and convenient basis for the set of all
+excited states.  Our construction of a basis for the set of excited
+states begins with the standard basis $\set{\ket{m}:m\in\A_n\p{N}}$
+for the PS manifold $\M_0$, where $m\equiv\p{m_1,m_2,\cdots,m_n}$ and
+$m_\mu$ is the (integer) occupation number of the single-spin state
+$\mu$.  Written in the ``computational basis'' of $N$ multilevel
+spins, these states are
+\begin{align}
+  \ket{m} =
+  \f1{\sqrt{\C\p{m}}}
+  \sum_{\substack{\t{distinct}\\\t{permutations}\\\Pi~\t{of}~\tilde m}}
+  \Pi \ket{\tilde m},
+  &&
+  \ket{\tilde m} \equiv \bigotimes_\mu \ket{\mu}^{\otimes m_\mu},
+  \label{eq:pops_expansion}
+\end{align}
+where $\C\p{m}$ is a normalization factor equal to the number of terms
+in the sum.  The basic idea in constructing a basis for the set of
+excited states is to find a set of operators that give us control over
+the signs of the terms in the expansion above, which is sufficient
+control to generate {\it any} state that is consistent with the
+populations $m$.  Framing the problem of generating excited states as
+one of controlling signs in \eqref{eq:pops_expansion} makes it clear
+that we can freely restrict ourselves to considering only operators
+that are $\pm1$ on the diagonal (and zero elsewhere) in the
+computational basis.
+
+The set of all operators that are $\pm1$ on the diagonal (in the
+computational basis) is spanned by sums and tensor products of the
+single-spin operators
+$Z_\nu\equiv1-2\op\nu=\sum_{\mu\ne\nu}\op\mu-\op\nu$.  Given a list of
+non-negative integers $z\equiv\p{z_1,z_2,\cdots,z_n}$ with
+$\sum_\mu z_\mu=r$, we thus define the $r$-spin operator
+\begin{align}
+  X_z \equiv \bigotimes_\mu Z_\mu^{z_\mu},
+\end{align}
+and nominally use states of the form
+$\ket{m,\m w,z}=\O\p{\m w,X_z}\ket{m}$ as our basis for the $r$-magnon
+manifold $\M_r$.  There are three points to address concerning the
+states $\ket{m,\m w,z}$:
+\begin{enumerate*}
+\item depending on the value of $m$, some choices of $z$ result in
+  trivial states, $\ket{m,\m w,z}=0$;
+  \label{pt:trivial}
+\item the set of nontrivial states is generally overcomplete, in that
+  any given state can be written as a nontrivial linear combination of
+  others, and finally
+  \label{pt:dependent}
+\item even after identifying a linearly independent set of nontrivial
+  states, any given pair of states $\ket{m,\m w,y},\ket{m,\m w,z}$
+  with $y\ne z$ generally have a non-zero overlap, which is to say
+  that they are not orthogonal.
+  \label{pt:orthogonal}
+\end{enumerate*}
+To address the first two points, \ref{pt:trivial} and
+\ref{pt:dependent}, we will place restrictions on the values of $z$
+that we allow for any given $m$.  The last point, \ref{pt:orthogonal},
+will require us to construct a basis of orthonormal states numerically
+for any given values of $m,\m w$.
+
+First, we identify criteria under which we are guaranteed that
+$\ket{m,\m w,z}=0$.  We begin by assuming that $\ket{m,\m w,z}\in\M_r$
+with $\sum_\nu z_\nu=r$.  We then observe that if $z_\nu>m_\nu$, then
+$m_\nu-z_\nu$ factors of $Z_\nu$ in $X_z$ are ``wasted,'' in the sense
+that they never yield relative signs in different terms of $\ket{m}$
+when expressed in the computational basis, as in
+\eqref{eq:pops_expansion}.  More concretely, consider the individual
+terms of the expansion
+\begin{align}
+  \ket{m,\m w,z}
+  = \f1{\sqrt{\C\p{m}}}
+  \sum_{\substack{\t{distinct}\\\t{permutations}\\\Pi~\t{of}~\tilde m}}
+  \sum_{k\in\D_N\p{r}} w_k \sp{X_z}_k \Pi \ket{\tilde m},
+  \label{eq:excited_state_expansion}
+\end{align}
+where $\sp{X_z}_k$ denotes the action of $X_z$ on spins
+$k=\p{k_1,k_2,\cdots,k_r}$.  All terms in
+\eqref{eq:excited_state_expansion} have exactly $m_\nu$ spins in the
+state $\nu$.  For any fixed $k$, at least $m_\nu-z_\nu$ factors of
+$Z_\nu$ in $\sp{X_z}_k$ must therefore evaluate to $+1$.  We can
+eliminate these ``extra'' factors of $Z_\nu$ by replacing
+$m_\nu-z_\nu$ of the factors that evaluate to $+1$ by the identity
+operator.  Performing this replacement for all $\nu\in\ZZ_n$ yields a
+state $\ket{m,\m w_{mz},{\floor{z}_m}}=\ket{m,\m w,z}$, where
+$\floor{z}_m\equiv\p{\min\set{z_\nu,m_\nu}:\nu\in\ZZ_n}$ indexes the
+$q$-spin operator $X_{\floor{z}_m}$, and $\m w_{mz}$ is some new
+tensor of coefficients.  If $q<r$ (because $z_\nu>m_\nu$ for some
+$\nu$), then our initial assumption that $\ket{m,\m w,z}\in\M_r$
+implies that $\ket{m,\m w,z}=0$.  In other words, a nontrivial basis
+$\set{\ket{m,\m w,z}}$ for $\M_r$ must have $z_\nu\le m_\nu$ for all
+$\nu$.
+
+Next, we address the fact that the basis $\set{\ket{m,\m w,z}}$ for
+$\M_r$ is overcomplete.  This fact is a simple consequence of our
+choice of single-spin operators for the construction of
+$\ket{m,\m w,z}$, namely $\set{Z_\nu:\nu\in\ZZ_n}$.  In essence, our
+problem is the fact that $Z_\nu=\p{n-2}-\sum_{\mu\ne\nu}Z_\mu$, which
+implies that any state $\ket{m,\m w,z}\in\M_r$ can be written as a
+linear combination of states in which $z_\nu\to z_\nu-1$, and possibly
+$z_\mu\to z_\mu+1$ for some $\mu$.  We can eliminate this redundancy
+by removing an operator from the $\set{Z_\nu}$, or equivalently
+enforcing that $z_\nu=0$ for some $\nu$.  As a matter of choice, we
+therefore consider only states $\ket{m,\m w,z}$ for which $z$ is zero
+at the first entry where $m$ is maximal: $z_\nu=0$ for
+$\nu=\min\set{\mu:m_\mu=\max\p{m}}$.  Given our previous restriction
+that $z_\nu\le m_\nu$ for all $\nu$, enforcing that $z$ is zero at an
+entry where $m$ is maximal maximizes the number of candidate values of
+$z$ that we eliminate from consideration.
+
+The above restrictions on $z$ do not {\it guarantee} that
+$\ket{m,\m w,z}\ne0$, or that the restricted basis
+$\set{\ket{m,\m w,z}}$ for $\M_r$ will not be overcomplete.
+Furthermore, this restricted basis is not orthogonal, which is to say
+that generally $\bk{m,\m v,y|m,\m w,z}\ne0$ even when $y\ne z$.  In
+order to construct an orthonormal basis for $\M_r$, we therefore need
+to build a matrix $B_m$ with matrix elements
+\begin{align}
+  B_m\p{\m v,y;\m w,z} \equiv \bk{m,\m v,y|m,\m w,z},
+\end{align}
+where $y,z$ satisfy the restrictions outlined above, and $\m v,\m w$
+are chosen from the set of tensors that generate eigenstates of $H_0$.
+Eigenvectors of $B_m$ are then mutually orthogonal vectors of the form
+$\sum_{\m w,z}c_{\m w z}\ket{m,\m w,z}$, whose squared norms are equal
+to their corresponding eigenvalues.  Note that the matrix $B_m$ has a
+block-diagonal structure, as it cannot (for example) couple states
+$\ket{m,\m v,y},\ket{m,\m w,z}$ whose tensors $\m v,\m w$ generate
+states with different interaction energies.  We therefore only need to
+build and diagonalize this matrix independently within each block.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Permutation-symmetrized products of multi-body operators}
+\section{Products of multi-body operators}
 \label{sec:operator_product}

### 2020-06-17 collective_sun/sun_perturbation.tex
amended some discussions about the operator content of energy eigenstates

@@
 Our starting basis for $\M_r$ then consists of states
-$\ket{m,\m w,z}\equiv\O\p{\m w,X_z}\ket{m}$, where we enforce that
+$\ket{m,\m w,X_z}\equiv\O\p{\m w,X_z}\ket{m}$, where we enforce that
 $z_\nu\le m_\nu$ for all $\nu$ (where $m_\nu$ is the occupation number
@@
 making the remaining set linearly independent of the identity
-operator.  The basis $\set{\ket{m,\m w,z}}$ for $\M_r$ is generally
-complete, but {\it not} orthonormal, so as a final step we construct
-an orthonormal basis $\set{\ket{m,\m w,X}}$, where for operators $X$
-we define $\ket{m,\m w,X}\equiv\O\p{\m w,X}\ket{m}$ (see Appendix
-\ref{sec:operator_content}).
+operator.  The basis $\set{\ket{m,\m w,X_z}}$ for $\M_r$ is generally
+complete, but not orthonormal, so as a final step we use the states
+$\ket{m,\m w,X_z}$ to (numerically) construct an orthonormal basis
+$\B_r$ for $\M_r$ (see Appendix \ref{sec:operator_content}).
 
@@
 
-The above restrictions on $z$ do not {\it guarantee} that
-$\ket{m,\m w,z}\ne0$, or that the restricted basis
-$\set{\ket{m,\m w,z}}$ for $\M_r$ will not be overcomplete.
-Furthermore, this restricted basis is not orthogonal, which is to say
-that generally $\bk{m,\m v,y|m,\m w,z}\ne0$ even when $y\ne z$.  In
-order to construct an orthonormal basis for $\M_r$, we therefore need
-to build a matrix $B_m$ with matrix elements
-\begin{align}
-  B_m\p{\m v,y;\m w,z} \equiv \bk{m,\m v,y|m,\m w,z},
-\end{align}
-where $y,z$ satisfy the restrictions outlined above, and $\m v,\m w$
-are chosen from the set of tensors that generate eigenstates of $H_0$.
+While the above restrictions on $z$ eliminate some trivial and
+redundant states from the basis $\set{\ket{m,\m w,z}}$ for $\M_r$,
+they do not guarantee that there will be {\it no} trivial states or
+redundancies left over.  Furthermore, even after restrictions the
+remaining basis is not orthogonal, which is to say that generally
+$\bk{m,\m v,y|m,\m w,z}\ne0$ even when $y\ne z$.  In order to
+construct an orthonormal basis for $\M_r$, we therefore need to build
+a matrix $B_m$ with matrix elements
+\begin{align}
+  B_m\p{\m v,y;\m w,z} \equiv \bk{m,\m v,y|m,\m w,z}
+  = \bk{m| \O\p{\m v,X_y}^\dag \O\p{\m w,X_z} |m},
+\end{align}
+where $y,z$ satisfy appropriate restrictions, and $\m v,\m w$ are
+chosen from the set of tensors that generate eigenstates of $H_0$.
 Eigenvectors of $B_m$ are then mutually orthogonal vectors of the form

### 2020-06-19 collective_sun/sun_perturbation.tex
minor wording changes

@@
 
+% focus on energy -- how different from non-interacting magnons?
+% plot dispersion relation
+% lax formulation for single-particle terms
+
 The label $m$ suffices to identify the PS (Dicke) manifold
@@
     as well as $\M_0$ itself, while $r$-body operators
-    $\O\p{\m w,\sigma_\z^{\otimes r}}$ couple $\M_0$ to all $q$-magnon
-    manifolds $\M_q$ with $q\le r$.}
+    $\O\p{\m w,\sigma_\z^{\otimes r}}$ generally couple $\M_0$ to all
+    $q$-magnon manifolds $\M_q$ with $q\le r$.}
   \label{fig:manifold_sketch}
@@
 of non-negative integers $z\equiv\p{z_1,z_2,\cdots,z_n}$ with
-$\sum_\mu z_\mu=r$ to define the $r$-spin operator
+$\sum_\nu z_\nu=r$ to define the $r$-spin operator
 \begin{align}
@@
 linear combination of states in which $z_\nu\to z_\nu-1$, and possibly
-$z_\mu\to z_\mu+1$ for some $\mu$.  We can eliminate this redundancy
-by removing an operator from the $\set{Z_\nu}$, or equivalently
-enforcing that $z_\nu=0$ for some $\nu$.  As a matter of choice, we
-therefore consider only states $\ket{m,\m w,z}$ for which $z$ is zero
-at the first entry where $m$ is maximal: $z_\nu=0$ for
+$z_\mu\to z_\mu+1$ for some $\mu\ne\nu$.  We can eliminate this
+redundancy by removing an operator from the $\set{Z_\nu}$, or
+equivalently enforcing that $z_\nu=0$ for some $\nu$.  As a matter of
+choice, we therefore consider only states $\ket{m,\m w,z}$ for which
+$z$ is zero at the first entry where $m$ is maximal: $z_\nu=0$ for
 $\nu=\min\set{\mu:m_\mu=\max\p{m}}$.  Given our previous restriction
@@
 for all $\alpha\in\ZZ_p$.  A classification of the terms in
-\eqref{eq:sym_prod_start} by diagrams $g\in\G\p{\m M}$ allows us to
+\eqref{eq:sym_prod_sum} by diagrams $g\in\G\p{\m M}$ allows us to
 expand
@@
 where $\EQPS$ denotes equality under a restriction to $\M_0$, and
-$\Q_g\p{\m w,\m X}$ is the sum of all terms in
-\eqref{eq:sym_prod_start} that are associated with the diagram $g$.
+$\Q_g\p{\m w,\m X}$ is the sum of all terms in \eqref{eq:sym_prod_sum}
+that are associated with the diagram $g$.
 

### 2020-06-22 collective_sun/sun_perturbation.tex
added text describing truncated shell simulations

@@
 generally the manifold of $r$-magnon states with $S=N/2-r$ is $\M_r$.
-Rather than the spin projection $m_\z$ we then identify states by
+Rather than the spin projection $m_\z$, we then identify states by
 labels $m=\p{m_\up,m_\dn}\in\A_2\p{N}$ in which $m_\mu$ denotes the
-occupation number of the single-spin state $\mu$, with
-$m_\up+m_\dn=N$ and $m_\up-m_\dn=m_\z$.
-
-\begin{figure}
-  \centering
-  \includegraphics{structure/mat_L101_a2.pdf}
-  \caption{Visualizations of dimension-$r$ tensors $\m w$ that
-    generate the lowest-lying (unnormalized) $r$-magnon states
-    $\ket{\m w,m} \equiv \O\p{\m w, \sigma_\z^{\otimes{r}}} \ket{m}$
-    with definite excitation energy $\Delta$ above the PS manifold
-    $\M_0$ in a periodic 1D lattice of $N=101$ spins interacting via
-    the SU(2)-symmetric Hamiltonian in \eqref{eq:H_SU2} with power-law
-    couplings $h_{ij}=-1/\oper{dist}\p{i,j}^2$.  Here
-    $\oper{dist}\p{i,j}\equiv\min\p{\abs{i-j},N-\abs{i-j}}$ is the
-    minimal distance between spins $i,j$.  Top row ($r=2$) shows the
-    full 2D matrix $\m w = \sum_{i,j}w_{ij}\op{i}{j}$, and bottom row
-    ($r=3$) shows a single 2D slice of the full 3D tensor,
-    $\tilde{\m w}_0 = \sum_{i,j}w_{ij,0}\op{i}{j}$.  Blue (red) pixels
-    correspond to matrix elements with a positive (negative) sign, and
-    darker colors indicate larger absolute values.  In the case of
-    $r=2$, translational invariance enforces $w_{ij}=w_{i-d,j-d}$ for
-    all displacements $d$, so every row of $\m w$ is some translation
-    of a single (arbitrary) row.  Similarly, when $r=3$ every 2D slice
-    of the full 3D tensor can be obtained by appropriately translating
-    the 2D slice $\tilde{\m w}_0$ shown above.  In the case of SU($n$)
-    with $n>2$, the ``$r$-magnon'' excitation energies $\Delta$ and
-    tensors $\m w$ remain identical, but now correspond to a
-    degenerate set of states that are generated by different choices
-    of $r$-body operators in place of $\sigma_\z^{\otimes r}$.}
-  \label{fig:structure}
-\end{figure}
+occupation number of the single-spin state $\mu$, with $m_\up+m_\dn=N$
+and $m_\up-m_\dn=m_\z$.
+
+% \begin{figure}
+%   \centering
+%   \includegraphics{structure/mat_L101_a2.pdf}
+%   \caption{Visualizations of dimension-$r$ tensors $\m w$ that
+%     generate the lowest-lying (unnormalized) $r$-magnon states
+%     $\ket{\m w,m} \equiv \O\p{\m w, \sigma_\z^{\otimes{r}}} \ket{m}$
+%     with definite excitation energy $\Delta$ above the PS manifold
+%     $\M_0$ in a periodic 1D lattice of $N=101$ spins interacting via
+%     the SU(2)-symmetric Hamiltonian in \eqref{eq:H_SU2} with power-law
+%     couplings $h_{ij}=-1/\oper{dist}\p{i,j}^2$.  Here
+%     $\oper{dist}\p{i,j}\equiv\min\p{\abs{i-j},N-\abs{i-j}}$ is the
+%     minimal distance between spins $i,j$.  Top row ($r=2$) shows the
+%     full 2D matrix $\m w = \sum_{i,j}w_{ij}\op{i}{j}$, and bottom row
+%     ($r=3$) shows a single 2D slice of the full 3D tensor,
+%     $\tilde{\m w}_0 = \sum_{i,j}w_{ij,0}\op{i}{j}$.  Blue (red) pixels
+%     correspond to matrix elements with a positive (negative) sign, and
+%     darker colors indicate larger absolute values.  In the case of
+%     $r=2$, translational invariance enforces $w_{ij}=w_{i-d,j-d}$ for
+%     all displacements $d$, so every row of $\m w$ is some translation
+%     of a single (arbitrary) row.  Similarly, when $r=3$ every 2D slice
+%     of the full 3D tensor can be obtained by appropriately translating
+%     the 2D slice $\tilde{\m w}_0$ shown above.  In the case of SU($n$)
+%     with $n>2$, the ``$r$-magnon'' excitation energies $\Delta$ and
+%     tensors $\m w$ remain identical, but now correspond to a
+%     degenerate set of states that are generated by different choices
+%     of $r$-body operators in place of $\sigma_\z^{\otimes r}$.}
+%   \label{fig:structure}
+% \end{figure}
 
@@
 \begin{align}
-  \ket{m,\m v} \equiv \O\p{\m v,\sigma_\z} \ket{m}
-  = \sum_j v_j \sigma_\z^{(j)} \ket{m},
+  \ket{m,\m v,\sigma_z}
+  \equiv \O\p{\m v,\sigma_\z} \ket{m}
+  = \sum_{j=1}^N v_j \sigma_\z^{(j)} \ket{m},
 \end{align}
@@
 $k_{\m v}\ne0$.  Similarly, $r$-magnon states for SU(2) take the form
-$\ket{m,\m w}\equiv\O\p{\m w,\sigma_\z^{\otimes r}}\ket{m}$, where now
-$\m w$ is an $r$-index tensor that encodes the structure of
-correlations in the excited state $\ket{m,\m w}$ (see Figure
-\ref{fig:structure}).  Considering states of this form, built using
-tensor products of Pauli-$z$ ($\sigma_\z$) operators, suffices to
-construct a basis for all excited states of an SU(2) spin system (see
-Appendix \ref{sec:operator_content}).  Note that $r$-body operators
-with arbitrary coefficients $\m w$ will generally couple the PS
-manifold $\M_0$ to all $q$-magnon manifolds $\M_q$ with $q\le r$ (see
-Figure \ref{fig:manifold_sketch}).  In order to generalize the notion
-of $r$-magnon manifolds $\M_r$ to the case of SU($n$) with $n>2$,
-where the analogous notion of a spin length $S$ is less intuitive, we
-therefore define $\M_r$ for $r>0$ as the manifold of states that (i)
-can be generated by the action of $r$-body operators on $\M_0$, and
-(ii) are orthogonal to all states in $\bigoplus_{q<r}\M_q$.
+$\ket{m,\m w,\sigma_\z^{\otimes r}} \equiv \O\p{\m
+  w,\sigma_\z^{\otimes r}}\ket{m}$, where now $\m w$ is an $r$-index
+tensor that encodes the structure of correlations in the excited state
+$\ket{m,\m w,\sigma_\z^{\otimes r}}$.  Considering states of this
+form, built using tensor products of Pauli-$z$ ($\sigma_\z$)
+operators, suffices to construct a basis for all excited states of an
+SU(2) spin system (see Appendix \ref{sec:operator_content}).  Note
+that $r$-body operators with arbitrary coefficients $\m w$ will
+generally couple the PS manifold $\M_0$ to all $q$-magnon manifolds
+$\M_q$ with $q\le r$ (see Figure \ref{fig:manifold_sketch}).  In order
+to generalize the notion of $r$-magnon manifolds $\M_r$ to the case of
+SU($n$) with $n>2$, where the analogous notion of a spin length $S$ is
+less intuitive, we can define $\M_r$ for $r>0$ as the manifold of
+states that (i) can be generated by the action of $r$-body operators
+on $\M_0$, and (ii) are orthogonal to all states in
+$\bigoplus_{q<r}\M_q$.
 
@@
 $\M_0=\oper{span}\set{\ket{m}:m\in\A_n\p{N}}$, and generate excited
-states by the action of multi-body operators, $\O\p{\m w,X}\ket{m}$.
-The choice of tensors $\m w$ for generating states in the $r$-magnon
-manifold $\M_r$ is the same as before, but we now have the additional
-complication of having to choose $r$-spin operators, $X$.  As with the
-choice of $\sigma_\z^{\otimes r}$ for SU(2), it suffices to choose
-tensor products of {\it diagonal} single-spin operators (that is,
-diagonal in a fixed single-particle basis), but there will now be
-$\sim\p{n-1}^r$ choices: one of $n-1$ diagonal operators (that are
-linearly independent of the identity) for each of $r$ tensor factors.
-We discuss the exact choice of operators in more detail in Appendix
+states by the action of multi-body operators,
+$\ket{m,\m w,X}\equiv\O\p{\m w,X}\ket{m}$.  The choice of tensors
+$\m w$ for generating states in the $r$-magnon manifold $\M_r$ is the
+same as before, but we now have the additional complication of having
+to choose $r$-spin operators, $X$.  As with the choice of
+$\sigma_\z^{\otimes r}$ for SU(2), it suffices to choose tensor
+products of {\it diagonal} single-spin operators (that is, diagonal in
+a fixed single-particle basis), but there will now be $\sim\p{n-1}^r$
+choices: one of $n-1$ diagonal operators (that are linearly
+independent of the identity) for each of $r$ tensor factors.  We
+discuss the exact choice of operators in more detail in Appendix
 \ref{sec:operator_content}.  To summarize, we define $n$ single-spin
@@
 Our starting basis for $\M_r$ then consists of states
-$\ket{m,\m w,X_z}\equiv\O\p{\m w,X_z}\ket{m}$, where we enforce that
-$z_\nu\le m_\nu$ for all $\nu$ (where $m_\nu$ is the occupation number
-of the single-spin state $\nu$), and that $z$ is zero in an entry
-where $m$ is maximal: $z_\nu=0$ for
-$\nu=\min\set{\mu:m_\mu=\max\p{m}}$.  This second restriction is
-equivalent to removing an operator from $\set{Z_\nu:\nu\in\ZZ_n}$,
-making the remaining set linearly independent of the identity
-operator.  The basis $\set{\ket{m,\m w,X_z}}$ for $\M_r$ is generally
-complete, but not orthonormal, so as a final step we use the states
-$\ket{m,\m w,X_z}$ to (numerically) construct an orthonormal basis
-$\B_r$ for $\M_r$ (see Appendix \ref{sec:operator_content}).
-
-\red{[todo: discuss the use of this basis in numerical simulations]}
+$\ket{m,\m w,X_z}$, where we enforce that $z_\nu\le m_\nu$ for all
+$\nu$ (where $m_\nu$ is the occupation number of the single-spin state
+$\nu$), and that $z$ is zero in an entry where $m$ is maximal:
+$z_\nu=0$ for $\nu=\min\set{\mu:m_\mu=\max\p{m}}$.  This second
+restriction is equivalent to removing an operator from
+$\set{Z_\nu:\nu\in\ZZ_n}$, making the remaining set linearly
+independent of the identity operator.  The basis
+$\set{\ket{m,\m w,X_z}}$ for $\M_r$ is generally complete, but not
+orthonormal, so as a final step we use the states $\ket{m,\m w,X_z}$
+to (numerically) construct an orthonormal basis $\B_r$ for $\M_r$ (see
+Appendix \ref{sec:operator_content}).
+
+By repeating the above process incrementally for all
+$r\le r_{\t{max}}$ (starting with $r=1$), we can construct an
+orthonormal basis
+$\col\B_{r_{\t{max}}}\equiv\bigotimes_{r\le r_{\t{max}}}\B_r$ for the
+combined manifold of all states with at most $r_{\t{max}}$ magnons,
+$\col\M_{r_{\t{max}}}\equiv\bigotimes_{r\le r_{\t{max}}}\M_r$.  The
+purpose of constructing the basis $\col\B_{r_{\t{max}}}$ is to
+simulate the dynamics induced by the addition of a spin operator
+$\O\p{\m v,Y}$ to the base Hamilonian $H_0$, such that the total
+Hamiltonian is $H = H_0 + \O\p{\m v,Y}$.  Throughout the construction
+of $\col\B_s$, an appropriate choice of tensors $\m w$ ensures that
+all states $\ket{m,\m w,X}\in\col\B_s$ are eigenstates of the base
+Hamiltonian $H_0$ (see Appendix \ref{sec:eigenstates}), which is to
+say that $H_0$ is diagonal with respect to $\col\B_s$.  Matrix
+elements of $\O\p{\m v,Y}$, meanwhile, take the form
+\begin{align}
+  \bk{\ell,\m u,X|\O\p{\m v,Y}|m,\m w,Z}
+  = \bk{\ell|\O\p{\m u,X}\O\p{\m v,Y}\O\p{\m w,Z}|m}.
+\end{align}
+In order to compute these matrix elements, we simplify general
+products of multi-body operators in Appendix
+\ref{sec:operator_product}.  For brevity, we will refer to numerical
+simulations restricted to $\col\M_{r_{\t{max}}}$ as {\it
+  order-$r_{\t{max}}$ truncated shell} (TS$_{r_{\t{max}}}$)
+simulations.
+
+An $M$-spin operator $\O\p{\m v,Y}$ can only couple $r$-magnon states
+to $q$-magnon states with $\abs{q-r}\le M$.  In the perturbative
+regime, $\norm{\O\p{\m v,Y}}<\Delta_{\t{gap}}/2$, TS$_{r_{\t{max}}}$
+simulations of initially PS (zero-magnon) states therefore capture all
+virtual processes through order $2\floor{r_{\t{max}}/M}+1$ in
+perturbation theory.  When $\norm{\O\p{\m v,Y}}\ge\Delta_{\t{gap}}/2$,
+perturbation theory formally breaks down, and the error incurred by
+neglecting all $s$-magnon states with $s>r_{\t{max}}$ is not clear a
+priori.  Nonetheless, TS$_{r_{\t{max}}}$ simulations can be
+benchmarked by examining the populations of all $r$-magnon manifolds
+$\M_r$ with $r\le r_{\t{max}}$.  As long as the maximal populations
+within large-$r$ manifolds $\M_r$ are small, one should expect these
+simulations to faithfully capture (at least qualitatively) the
+dynamical behavior of the spin system under consideration.
 
 \newpage
-
-\red{[note: the remainder of this section is old, and will be replaced
-  entirely]}
-
-By construction, the $M$-body operator $X_{\m w}$ generally couples PS
-states in $\M_0$ to asymmetric states in $\col\M_M$.  If the coupling
-between $\M_0$ and its complement is weak compared to the spectral gap
-$\Delta_{\t{gap}}$ of $H_0$, then we can treat the effect of
-$X_{\m w}$ perturbatively and derive an effective Hamiltonian
-$H_{\t{eff}}$ on $\M_0$.  This ground-state perturbation theory
-formally breaks down when the operator norm
-$\norm{X_{\m w}}\ge\Delta_{\t{gap}}/2$, in which case the operator
-$X_{\m w}$ first couples $\M_0$ to $\col\M_M$, then $\col\M_M$ to
-$\col\M_{2M}$, and so on.  Even if
-$\norm{X_{\m w}}>\Delta_{\t{gap}}/2$, however, the energetic penalty
-incurred by $H_0$ from breaking the permutational symmetry of an
-initial state $\ket\psi\in\M_0$ suppresses the leakage of population
-into manifolds $\M_r$ of increasing $r$.  One might therefore capture
-the dynamics of an initial PS state $\ket\psi\in\M_0$ by projecting
-$X_{\m w}$ onto $\col\M_{r_{\t{max}}}$ for some $r_{\t{max}}\ge M$,
-resulting in a theory that resembles first-order perturbation theory
-on $\col\M_{r_{\t{max}}}$.  In the perturbative regime
-$\norm{X_{\m w}}<\Delta_{\t{gap}}/2$, such a theory captures all
-virtual processes through order $2\floor{r_{\t{max}}/M}+1$ in a formal
-perturbation theory on $\M_0$.  In the non-perturbative regime
-$\norm{X_{\m w}}\ge\Delta_{\t{gap}}/2$, the error incurred by
-truncating the Hilbert space of a spin system past
-$\col\M_{r_{\t{max}}}$ is not clear a priori.  Nonetheless,
-simulations restricted to $\col\M_{r_{\t{max}}}$ can be benchmarked by
-examining the maximum populations within all $\M_r$ for
-$r\le r_{\t{max}}$ throughout simulation.  As long as
-\begin{enumerate*}
-\item these populations fall off with increasing $r$,
-\item and the maximal population within large-$r$ manifolds $\M_r$ is
-  small,
-\end{enumerate*}
-one should expect such simulations to faithfully capture (at least
-qualitatively) the dynamical behavior of the spin system under
-consideration.
-
-In order to project an operator $X_{\m w}$ onto
-$\col\M_{r_{\t{max}}}$, we first need to construct a basis for
-$\col\M_{r_{\t{max}}}$, and then compute all matrix elements of
-$X_{\m w}$ in this basis.  Our strategy will be to first identify a
-basis for $\M_0$, and then find, for each $r=1,\cdots,r_{\t{max}}$, a
-suitable set of $r$-body operators $\set{O}$ and dimension-$r$ tensors
-$\set{\m u}$ for which the states
-$\ket{O,\m u,m}\propto O_{\m u}\ket{m}$ form an orthonormal basis for
-$\M_r$.  Matrix elements of $X_{\m w}$ with respect to this basis for
-$\col\M_{r_{\t{max}}}$ then take the form
-\begin{align}
-  \bk{O,\m u,\ell|X_{\m w}|Q,\m v,m}
-  = \f{\bk{\ell|O_{\m u}^\dag X_{\m w} Q_{\m v}|m}}
-  {\sqrt{\bk{\ell|O_{\m u}^\dag O_{\m u}|\ell}
-      \bk{m|Q_{\m v}^\dag Q_{\m v}|m}}}.
-  \label{eq:matrix_element}
-\end{align}
-In order to compute these matrix elements, we simplify products of the
-form $\P_0 O_{\m u}^\dag X_{\m w} Q_{\m v} \P_0$ and
-$\P_0 O_{\m u}^\dag O_{\m u} \P_0$, with $\P_0$ a projector onto
-$\M_0$, in Appendix \ref{sec:operator_product}.
-
-Denoting all assignments of $N$ (identical) spins to $n$ (distinct)
-states by $\A_n\p{N}$, a suitable basis for $\M_0$ is the set of PS
-states $\set{\ket{m}:m\in\A_n\p{N}}$, where $m=\p{m_1,m_2,\cdots,m_n}$
-specifies the occupation number $m_\mu$ of each single-spin state
-$\mu$.  For consistency with the expression of matrix elements in
-\eqref{eq:matrix_element}, for each $\ket{m}\in\M_0$ we define
-$\ket{\1,1,m}\equiv\ket{m}$, with $\1$ the identity operator, and we
-define $\1_1\equiv\1$.  We now need to determine the set of $r$-body
-operators $\set{O}$ and dimension-$r$ tensors $\set{\m u}$ that define
-our basis for each $\M_r$.  As an initial guess for $\set{\m u}$, we
-can choose the set $\set{\m u_\Delta}_{\m h,r}$ of dimension-$r$
-tensors $\m u_\Delta$ that generate eigenstates of $H_0$.  The set of
-$r$-body operators to choose is less obvious, but can be determined by
-an exhaustive search over all $r$-body operators relevant to a
-particular scenario.  Specifically, the purpose of projecting
-operators onto $\col\M_{r_{\t{max}}}$ is to simulate the dynamics
-induced by some operator $\hat\O$ of the form in \eqref{eq:pert},
-constructed from a set $\O=\set{\p{X,\m w}}$ containing multi-spin
-operators $X$.  In order to find a basis for the {\it relevant} states
-in $\M_r$, one can simply consider the set $\set{O}_{\O,r}$ all
-$r$-spin operators that can be constructed from products of the
-multi-spin operators $X$ in $\O$.
-
-In general, the states $\ket{O,\m u_\Delta,m}$ for all
-$\p{O,\m u_\Delta,m}\in\set{O}_{\O,r}\times\set{\m u_\Delta}_{\m h,r}
-\times\A_n\p{N}$ will form an overcomplete basis for the relevant
-states in $\col\M_r$.  Our last step is therefore to reduce this set
-of states to a minimal basis $\B_r$ for $\M_r$.  To this end, we
-construct each of $\B_1,\B_2,\cdots,\B_{r_{\t{max}}}$ in sequence,
-such that when constructing $\B_r$ we already have a minimal basis
-$\col\B_{r-1}\equiv\bigcup_{s<r}\B_s$ for $\col\M_{r-1}$.  To
-construct $\B_r$, we then consider each
-$\p{O,\m u_\Delta,m}\in\set{O}_{\O,r}\times\set{\m u_\Delta}_{\m h,r}
-\times\A_n\p{N}$ one at a time.  If $\ket{O,\m u_\Delta,m}$ is
-orthogonal to all states in $\col\B_{r-1}$, as well as all other
-states we have added to $\B_r$ thus far, then we include it in $\B_r$.
-Otherwise, we discard $\ket{O,\m u_\Delta,m}$.  There are two
-considerations to keep in mind for this construction: first, one
-should always throw out null states $\ket{O,\m u_\Delta,m}=0$, which
-can be identified by a vanishing self-overlap,
-$\bk{O,\m u_\Delta,m|O,\m u_\Delta,m}=0$.  Second, computing overlaps
-to check for orthogonality can be computationally intensive.  When
-considering a candidate state $\ket{O,\m u_\Delta,m}$ to add to
-$\B_r$, one can therefore restrict orthogonality checks to other
-states $\ket{Q,\m v_\epsilon,\ell}$ with $\epsilon=\Delta$; if the
-excitation energies $\epsilon\ne\Delta$, then these states are
-guaranteed to be orthogonal.  Any two states with different occupation
-numbers of single-spin states $\mu\in\ZZ_n$ are similarly guaranteed
-to be orthogonal.
-
-\red{[look at \cite{rai1988correlated} for classifying asymmetric
-  states?]}
 
@@
 with $m_{ii}=0$.  This product appears, for example, in the projection
-of Ising interactions onto the two-body excitation manifold $\M_2$.
-Collecting terms according to the number of $Z$ operators that remain
-after accounting for $Z^2=1$, we find that
+of Ising interactions onto the two-magnon manifold $\M_2$.  Collecting
+terms according to the number of $Z$ operators that remain after
+accounting for $Z^2=1$, we find that
 \begin{align}

### 2020-06-23 collective_sun/sun_perturbation.tex
rearranged discussion of truncated shell simulations added an intro for the appendix on multiplying multi-body operators

@@
 \begin{document}
-
-% todo: solve collective model; Casmir operators?
 
@@
 \begin{enumerate*}
-\item an initial state in the manifold of permutationally-symmetric
+\item an initial state in the manifold of permutationally symmetric
   states, a multilevel generalization of the Dicke manifold for
@@
 notes, but note that this restriction can be relaxed to the assumption
-that the initial state of all spins is a permutationally-symmetric
+that the initial state of all spins is a permutationally symmetric
 state (e.g.~a spin-polarized state).  We will also assume that the PS
@@
   = \sum_\Delta \p{E_0+\Delta} \O\p{\m w_\Delta,X} \ket\psi,
+  \label{eq:excitation}
 \end{align}
@@
 than treating these excited states as ``intermediate'' or ``virtual''
-in the derivation of an effective ground-state Hamiltonian, we can
+for the derivation of an effective ground-state Hamiltonian, we can
 perform numerical simulations in the joint Hilbert space of both
@@
 $\norm{\O\p{\m w,X}}<\Delta_{\t{gap}}/2$, such simulations are
-equivalent (up to vanishing perturbative corrections) to a
-higher-order perturbation theory.  Moreover, the first-class treatment
-of excited states in finite systems enables simulating the dynamics
-induced by {\it non-perturbative} additions $\O\p{\m w,X}$ to the base
+equivalent (up to perturbative corrections) to a higher-order
+perturbation theory.  Moreover, the first-class treatment of excited
+states in finite systems enables simulating the dynamics induced by
+{\it non-perturbative} additions $\O\p{\m w,X}$ to the base
 Hamiltonian $H_0$.
 
-In order to simulate dynamics within excited-state manifolds, we first
-need to classify the set of excited states, and construct a suitable
-basis.  In the case of SU(2), a common starting point is to identify
-states with a definite (non-negative) spin length
-$S=\set{N/2,N/2-1,\cdots}$, and a spin projection
-$m_\z\in\set{S,S-1,\cdots,-S}$ onto a quantization axis.  To
-generalize this classification to SU($n$), we denote the manifold of
-states with definite fixed spin length $S$ by $\M_{N/2-S}$.  The PS
-manifold of Dicke states with $S=N/2$ is then $\M_0$, the manifold of
-single-magnon spin-wave states with $S=N/2-1$ is $\M_1$, and more
+The basic idea will be to identify the set of states (manifolds)
+$\M_r$ with a definite number ($r$) of spin-wave excitations, or {\it
+  magnons}.  The ground-state manifold of PS states is simply the
+zero-magnon manifold $\M_0$.  If we construct a basis for all $\M_r$
+with $r\le r_{\t{max}}$, we can perform numerical simulations in the
+combined manifold of all states with at most $r_{\t{max}}$ magnons; we
+refer to such simulations as {\it order-}$r_{\t{max}}$ {\it truncated
+  shell} (TS$_{r_{\t{max}}}$) simulations.  An $M$-spin operator
+$\O\p{\m w,X}$ can only couple $r$-magnon states (in $\M_r$) to
+$q$-magnon states (in $\M_q$) with $\abs{q-r}\le M$.  For initially PS
+states, TS$_{r_{\t{max}}}$ simulations thus capture all virtual
+processes through order $2\floor{r_{\t{max}}/M}+1$ in perturbation
+theory, when $\norm{\O\p{\m w,X}}<\Delta_{\t{gap}}/2$.  When
+$\norm{\O\p{\m w,X}}\ge\Delta_{\t{gap}}/2$, perturbation theory
+formally breaks down, and the error incurred by neglecting all
+$s$-magnon states with $s>r_{\t{max}}$ is not clear a priori.
+Nonetheless, TS$_{r_{\t{max}}}$ simulations can be benchmarked by
+examining the populations of all $r$-magnon manifolds $\M_r$ with
+$r\le r_{\t{max}}$.  As long as the maximal populations within
+large-$r$ manifolds $\M_r$ are small, one should expect
+TS$_{r_{\t{max}}}$ simulations to faithfully capture (at least
+qualitatively) the dynamical behavior of the spin system under
+consideration.
+
+The remainder of this section is devoted to the construction of a
+basis for the $r$-magnon manifolds $\M_r$, and to the expansion of
+general spin operators $\O\p{\m w,X}$ in this basis.  We begin with a
+classification of excited states in the case of SU(2).  A common
+starting point is to first classify states by a definite
+(non-negative) spin length $S=\set{N/2,N/2-1,\cdots}$, and a spin
+projection $m_\z\in\set{S,S-1,\cdots,-S}$ onto a fixed quantization
+axis.  In fact, the manifold of states with definite fixed spin length
+$S$ is precisely the $\p{N/2-S}$-magnon manifold $\M_{N/2-S}$.  The PS
+manifold of Dicke states with $S=N/2$ is then $\M_0$, and more
 generally the manifold of $r$-magnon states with $S=N/2-r$ is $\M_r$.
-Rather than the spin projection $m_\z$, we then identify states by
-labels $m=\p{m_\up,m_\dn}\in\A_2\p{N}$ in which $m_\mu$ denotes the
-occupation number of the single-spin state $\mu$, with $m_\up+m_\dn=N$
-and $m_\up-m_\dn=m_\z$.
-
-% \begin{figure}
-%   \centering
-%   \includegraphics{structure/mat_L101_a2.pdf}
-%   \caption{Visualizations of dimension-$r$ tensors $\m w$ that
-%     generate the lowest-lying (unnormalized) $r$-magnon states
-%     $\ket{\m w,m} \equiv \O\p{\m w, \sigma_\z^{\otimes{r}}} \ket{m}$
-%     with definite excitation energy $\Delta$ above the PS manifold
-%     $\M_0$ in a periodic 1D lattice of $N=101$ spins interacting via
-%     the SU(2)-symmetric Hamiltonian in \eqref{eq:H_SU2} with power-law
-%     couplings $h_{ij}=-1/\oper{dist}\p{i,j}^2$.  Here
-%     $\oper{dist}\p{i,j}\equiv\min\p{\abs{i-j},N-\abs{i-j}}$ is the
-%     minimal distance between spins $i,j$.  Top row ($r=2$) shows the
-%     full 2D matrix $\m w = \sum_{i,j}w_{ij}\op{i}{j}$, and bottom row
-%     ($r=3$) shows a single 2D slice of the full 3D tensor,
-%     $\tilde{\m w}_0 = \sum_{i,j}w_{ij,0}\op{i}{j}$.  Blue (red) pixels
-%     correspond to matrix elements with a positive (negative) sign, and
-%     darker colors indicate larger absolute values.  In the case of
-%     $r=2$, translational invariance enforces $w_{ij}=w_{i-d,j-d}$ for
-%     all displacements $d$, so every row of $\m w$ is some translation
-%     of a single (arbitrary) row.  Similarly, when $r=3$ every 2D slice
-%     of the full 3D tensor can be obtained by appropriately translating
-%     the 2D slice $\tilde{\m w}_0$ shown above.  In the case of SU($n$)
-%     with $n>2$, the ``$r$-magnon'' excitation energies $\Delta$ and
-%     tensors $\m w$ remain identical, but now correspond to a
-%     degenerate set of states that are generated by different choices
-%     of $r$-body operators in place of $\sigma_\z^{\otimes r}$.}
-%   \label{fig:structure}
-% \end{figure}
-
-% focus on energy -- how different from non-interacting magnons?
-% plot dispersion relation
-% lax formulation for single-particle terms
+Rather than the spin projection $m_\z$, we then identify states by the
+occupation numbers $m=\p{m_\up,m_\dn}\in\A_2\p{N}$, where $m_\mu$
+denotes the total number of individual spins in state $\mu$, with
+$m_\up+m_\dn=N$ and $m_\up-m_\dn=m_\z$.
 
@@
 where each $v_j$ is a scalar and $\sigma_\z$ is a Pauli-$z$ operator.
-If an SU(2) spin system is translationally invariant, then these
-spin-wave eigenstates have $v_j=e^{\i k_{\m v}\c j}$ with wavenumber
+If an SU(2) spin system is translationally invariant, then spin-wave
+eigenstates have $v_j=e^{\i k_{\m v}\c j}$ with wavenumber
 $k_{\m v}\ne0$.  Similarly, $r$-magnon states for SU(2) take the form
@@
 tensor that encodes the structure of correlations in the excited state
-$\ket{m,\m w,\sigma_\z^{\otimes r}}$.  Considering states of this
-form, built using tensor products of Pauli-$z$ ($\sigma_\z$)
-operators, suffices to construct a basis for all excited states of an
-SU(2) spin system (see Appendix \ref{sec:operator_content}).  Note
-that $r$-body operators with arbitrary coefficients $\m w$ will
-generally couple the PS manifold $\M_0$ to all $q$-magnon manifolds
-$\M_q$ with $q\le r$ (see Figure \ref{fig:manifold_sketch}).  In order
-to generalize the notion of $r$-magnon manifolds $\M_r$ to the case of
-SU($n$) with $n>2$, where the analogous notion of a spin length $S$ is
-less intuitive, we can define $\M_r$ for $r>0$ as the manifold of
-states that (i) can be generated by the action of $r$-body operators
-on $\M_0$, and (ii) are orthogonal to all states in
-$\bigoplus_{q<r}\M_q$.
+$\ket{m,\m w,\sigma_\z^{\otimes r}}$.  The coefficient tensor $\m w$
+determines the energy of the excited state
+$\ket{m,\m w,\sigma_\z^{\otimes r}}$ through the expansion in
+\eqref{eq:excitation}.  Considering states of this form, built using
+tensor products of Pauli-$z$ ($\sigma_\z$) operators, suffices to
+construct a basis for all excited states of an SU(2) spin system (see
+Appendix \ref{sec:operator_content}).  Note that $r$-body operators
+with arbitrary coefficients $\m w$ will generally couple the PS
+manifold $\M_0$ to all $q$-magnon manifolds $\M_q$ with $q\le r$ (see
+Figure \ref{fig:manifold_sketch}).  In order to generalize the notion
+of $r$-magnon manifolds $\M_r$ to the case of SU($n$) with $n>2$,
+where the analogous notion of a spin length $S$ is less intuitive, we
+can define $\M_r$ for $r>0$ as the manifold of states that (i) can be
+generated by the action of $r$-body operators on $\M_0$, and (ii) are
+orthogonal to all states in $\bigoplus_{q<r}\M_q$.
 
@@
 $\col\B_{r_{\t{max}}}\equiv\bigotimes_{r\le r_{\t{max}}}\B_r$ for the
-combined manifold of all states with at most $r_{\t{max}}$ magnons,
-$\col\M_{r_{\t{max}}}\equiv\bigotimes_{r\le r_{\t{max}}}\M_r$.  The
-purpose of constructing the basis $\col\B_{r_{\t{max}}}$ is to
-simulate the dynamics induced by the addition of a spin operator
-$\O\p{\m v,Y}$ to the base Hamilonian $H_0$, such that the total
-Hamiltonian is $H = H_0 + \O\p{\m v,Y}$.  Throughout the construction
-of $\col\B_s$, an appropriate choice of tensors $\m w$ ensures that
-all states $\ket{m,\m w,X}\in\col\B_s$ are eigenstates of the base
-Hamiltonian $H_0$ (see Appendix \ref{sec:eigenstates}), which is to
-say that $H_0$ is diagonal with respect to $\col\B_s$.  Matrix
-elements of $\O\p{\m v,Y}$, meanwhile, take the form
+combined manifold of all states with at most $r_{\t{max}}$ magnons.
+During the construction of $\col\B_s$, an appropriate choice of
+tensors $\m w$ ensures that all states $\ket{m,\m w,X}\in\col\B_s$ are
+eigenstates of the base Hamiltonian $H_0$ (see Appendix
+\ref{sec:eigenstates}), which is to say that $H_0$ is diagonal with
+respect to $\col\B_s$.  Matrix elements of general spin operators
+$\O\p{\m v,Y}$, meanwhile, take the form
 \begin{align}
@@
   = \bk{\ell|\O\p{\m u,X}\O\p{\m v,Y}\O\p{\m w,Z}|m}.
+  \label{eq:matrix_element}
 \end{align}
@@
 products of multi-body operators in Appendix
-\ref{sec:operator_product}.  For brevity, we will refer to numerical
-simulations restricted to $\col\M_{r_{\t{max}}}$ as {\it
-  order-$r_{\t{max}}$ truncated shell} (TS$_{r_{\t{max}}}$)
-simulations.
-
-An $M$-spin operator $\O\p{\m v,Y}$ can only couple $r$-magnon states
-to $q$-magnon states with $\abs{q-r}\le M$.  In the perturbative
-regime, $\norm{\O\p{\m v,Y}}<\Delta_{\t{gap}}/2$, TS$_{r_{\t{max}}}$
-simulations of initially PS (zero-magnon) states therefore capture all
-virtual processes through order $2\floor{r_{\t{max}}/M}+1$ in
-perturbation theory.  When $\norm{\O\p{\m v,Y}}\ge\Delta_{\t{gap}}/2$,
-perturbation theory formally breaks down, and the error incurred by
-neglecting all $s$-magnon states with $s>r_{\t{max}}$ is not clear a
-priori.  Nonetheless, TS$_{r_{\t{max}}}$ simulations can be
-benchmarked by examining the populations of all $r$-magnon manifolds
-$\M_r$ with $r\le r_{\t{max}}$.  As long as the maximal populations
-within large-$r$ manifolds $\M_r$ are small, one should expect these
-simulations to faithfully capture (at least qualitatively) the
-dynamical behavior of the spin system under consideration.
+\ref{sec:operator_product}.
+
+% dispersion relation? magnon interactions?
+% lax formulation for single-particle terms
+% solve collective model? Casmir operators?
 
@@
   = \bk{m| \O\p{\m v,X_y}^\dag \O\p{\m w,X_z} |m},
+  \label{eq:basis_matrix}
 \end{align}
@@
 
-\red{[TODO: write introductory text for this section]}
+The construction of a basis $\B_r$ for the $r$-magnon manifold $\M_r$
+requires multiplying two multi-body operators
+$\O\p{\m v,X_y}^\dag, \O\p{\m w,X_z}$, and projecting the result onto
+the permutationally symmetric (PS) manifold $\M_0$, as in
+\eqref{eq:basis_matrix}.  Similarly, computing matrix elements of
+operators $\O\p{\m v,Y}$ with respect to the basis
+$\col\B_{r_{\t{max}}}\equiv\bigotimes_{r\le r_{\t{max}}}\B_r$ requires
+multiplying three multi-body operators and projecting the result onto
+$\M_0$, as in \eqref{eq:matrix_element}.  In this section, we develop
+a diagrammatic technique for the sorts of calculations, and work out
+some special cases of common interest.
 
@@
 Given a system of $N$ spins, we wish to project a product of
-multi-body operators onto the permutationally symmetric (PS) manifold
-$\M_0$.  For $p$ multi-body operators, such a product takes the form
+multi-body operators onto the PS manifold.  For $p$ multi-body
+operators, such a product takes the form
 \begin{align}

### 2020-06-25 collective_sun/sun_perturbation.tex
clarified reason for choosing diagonal +/-1 operators for generating excited states

@@
 occupation numbers $m=\p{m_\up,m_\dn}\in\A_2\p{N}$, where $m_\mu$
-denotes the total number of individual spins in state $\mu$, with
+denotes the total number of spins in the state $\mu$, with
 $m_\up+m_\dn=N$ and $m_\up-m_\dn=m_\z$.
@@
 where $\C\p{m}$ is a normalization factor equal to the number of terms
-in the sum.  The basic idea in constructing a basis for the set of
-excited states is to find a set of operators that give us control over
-the signs of the terms in the expansion above, which is sufficient
-control to generate {\it any} state that is consistent with the
-populations $m$.  Framing the problem of generating excited states as
-one of controlling signs in \eqref{eq:pops_expansion} makes it clear
-that we can freely restrict ourselves to considering only operators
-that are $\pm1$ on the diagonal (and zero elsewhere) in the
-computational basis.
-
-The set of all operators that are $\pm1$ on the diagonal (in the
-computational basis) is spanned by sums and tensor products of the
-single-spin operators
+in the sum.  While $\ket{m}$ is the only PS state with occupation
+numbers $m$, the set of {\it all} states with occupation numbers $m$
+is spanned by states of the form
+\begin{align}
+  \ket{m,s} = \f1{\sqrt{\C\p{m}}}
+  \sum_{\substack{\t{distinct}\\\t{permutations}\\\Pi~\t{of}~\tilde m}}
+  s\p{\Pi} \Pi \ket{\tilde m},
+  \label{eq:pops_states}
+\end{align}
+where $s\p{\Pi}\in\set{\pm1}$ for all permutations $\Pi$.  As a
+consequence, when generating excited states of the form
+$\O\p{\m w,X}\ket{m}$, we can freely restrict ourselves to operators
+$X$ that only change the signs of individual terms (associated with
+permutations $\Pi$) in \eqref{eq:pops_expansion}, which suffices to
+construct any state $\ket{m,s}$ in \eqref{eq:pops_states}.  We
+therefore consider single-operators $X$ that have $\pm1$ on the
+diagonal (in the computational basis), and zero everywhere else.
+
+The set of all operators that are $\pm1$ on the diagonal is spanned by
+sums and tensor products of the single-spin operators
 $Z_\nu\equiv1-2\op\nu=\sum_{\mu\ne\nu}\op\mu-\op\nu$.  Given a list of

### 2020-06-25 collective_sun/multilevel_spin_notes.tex
added more information about raman drive

@@
 
-\newcommand{\FS}{\text{FS}}
+\newcommand{\PS}{\text{PS}}
+\newcommand{\col}{\underline}
 
@@
 $S_{\mu\nu}^{(p)}$ in terms of Pauli operators in order to convert
-$H_{\t{int}}$ into an SU(2) spin Hamiltonian\cite{he2019engineering}.
+$H_{\t{int}}$ into an SU(2) spin Hamiltonian \cite{he2019engineering}.
 To generalize this procedure to the case of SU($n$), we first define a
@@
 The most obvious generalization of Pauli operators to the case of
-SU($n$) are the {\it generalized Gell-Mann (GGM)
-  operators}\cite{hioe1981level, bertlmann2008bloch}, defined for
+SU($n$) are the {\it generalized Gell-Mann (GGM) operators}
+\cite{hioe1981level, bertlmann2008bloch}, defined for
 $j,k,\ell\in\set{0,1,\cdots,n-1}$ with $j<k$ and $\ell>0$ by
@@
 instead obeys the symmetries of the {\it polarization} or {\it
-  transition operators} defined by\cite{kryszewski2006alternative,
+  transition operators} defined by \cite{kryszewski2006alternative,
   bertlmann2008bloch}
@@
 is a collective SU(2) spin-$\alpha$ operator for $\p{n-1}$ 2-level
-spins\cite{perlin2020shorttime}.  More concretely, we can embed the
+spins \cite{perlin2020shorttime}.  More concretely, we can embed the
 Hilbert space of a single $n$-level spin into the
@@
 is a scale factor that will appear frequently in the theory of drive
-operators.
+operators.  Some other useful relations between drive and spin
+operators are
+\begin{align}
+  \xi_2 D_{2,0} \to -\f12\p{S_\x^2 + S_\y^2} + \f{n^2-1}{12},
+  \label{eq:spin_ops_2_0}
+\end{align}
+\begin{align}
+  \sqrt{3}\, \xi_2 D_{2,2} \to \f12\p{S_\x^2 - S_\y^2},
+  &&
+  \sqrt{3}\, \xi_2 D_{2,1} \to \f12\p{S_\z S_\x + S_\x S_\z}.
+  \label{eq:spin_ops_2_12}
+\end{align}
 
@@
 
-Weyl operators\cite{bertlmann2008bloch} and self-adjoint linear
-combinations thereof\cite{asadian2016heisenbergweyl} also provide an
+Weyl operators \cite{bertlmann2008bloch} and self-adjoint linear
+combinations thereof \cite{asadian2016heisenbergweyl} also provide an
 interesting basis of operators for finite-dimensional Hilbert spaces.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Subsystem permutations and the fully symmetric manifold}
+\subsection{Subsystem permutations and the permutationally symmetric
+  manifold}
 \label{sec:perm_ops}
@@
 $H_{\t{int}}$ makes it evident that its ground-state manifold
-$\M_{\FS}$ is the set of {\it fully symmetric} (FS) states that have
-eigenvalue 1 with respect to all pairwise permutation operators
-$\Pi_{pq}$.  The fully symmetric manifold $\M_\FS$ is spanned by a
-basis $\set{\ket{\v m}}$ of fully symmetric states that can be
-uniquely identified by a list of integers
-$\v m\equiv\p{m_0,m_1,\cdots,m_{n-1}}\in\ZZ_{N+1}^n$, where $m_s$
-indicates the total occupation number of subsystem state $s$, and
-$\sum_sm_s=N$.  In the case of $n=2$, the fully symmetric manifold
-$\M_\FS$ is precisely the Dicke manifold\cite{dicke1954coherence} of
+$\M_{\PS}$ is the set of {\it permutationally symmetric} (PS) states
+that have eigenvalue 1 with respect to all pairwise permutation
+operators $\Pi_{pq}$.  The PS manifold $\M_\PS$ is spanned by a basis
+$\set{\ket{\v m}}$ of PS states that can be uniquely identified by a
+list of integers $\v m\equiv\p{m_0,m_1,\cdots,m_{n-1}}\in\ZZ_{N+1}^n$,
+where $m_s$ indicates the total occupation number of subsystem state
+$s$, and $\sum_sm_s=N$.  In the case of $n=2$, the PS manifold
+$\M_\PS$ is precisely the Dicke manifold \cite{dicke1954coherence} of
 $N$ qubits, spanned by Dicke states $\set{\ket{m_0,m_1}}$ for which
@@
 occupation number of qubit state $\ket{0}$ ($\ket{1}$).  The dimension
-of fully symmetric manifold $\M_\FS$ is determined by the number of
-ways to assign $N$ subsystems to $n$ states, which is
-${N+n-1 \choose n-1}$.  The energy of any fully symmetric state
-$\ket{\psi_\FS}\in\M_\FS$ with respect to the interaction Hamiltonian
-$H_{\t{int}}$ in \eqref{eq:H_int_perm} is
+of PS manifold $\M_\PS$ is determined by the number of ways to assign
+$N$ subsystems to $n$ states, which is ${N+n-1 \choose n-1}$.  The
+energy of any PS state $\ket{\psi_\PS}\in\M_\PS$ with respect to the
+interaction Hamiltonian $H_{\t{int}}$ in \eqref{eq:H_int_perm} is
 $-u{N \choose 2}=-\p{1/2}N\p{N-1}u$.
 
-In addition to the fact that the fully symmetric manifold $\M_\FS$ is
-ground-state manifold of the interaction Hamltonian $H_{\t{int}}$, an
-important feature of $\M_\FS$ is its closure under the action of
-collective operators of the form $\Q=\sum_pQ^{(p)}$, as well as linear
-combinations and products thereof.  Closure of $\M_\FS$ under
+In addition to the fact that the PS manifold $\M_\PS$ is ground-state
+manifold of the interaction Hamltonian $H_{\t{int}}$, an important
+feature of $\M_\PS$ is its closure under the action of collective
+operators of the form $\Q=\sum_pQ^{(p)}$, as well as linear
+combinations and products thereof.  Closure of $\M_\PS$ under
 collective the action of collective operators, and in particular under
@@
 The single-particle Hamiltonian $H_{\t{lat}}^{(\phi)}$ breaks
-permutational symmetry, thereby coupling states within the fully
-symmetric manifold $\M_\FS$ to states outsides it.  If SOC is
-sufficiently weak, however, transitions outside the fully symmetric
-manifold $\M_\FS$ are energetically suppressed by interactions, and we
-can account for the effect of SOC perturbatively.  In the regime of
-weak SOC with $\abs{\phi}\ll1$, by considering
+permutational symmetry, thereby coupling states within the PS manifold
+$\M_\PS$ to states outsides it.  If SOC is sufficiently weak, however,
+transitions outside the PS manifold $\M_\PS$ are energetically
+suppressed by interactions, and we can account for the effect of SOC
+perturbatively.  In the regime of weak SOC with $\abs{\phi}\ll1$, by
+considering
 \begin{enumerate*}
@@
 $\phi$ yields the following effective Hamiltonian in the fully
-symmetric manifold $\M_\FS$ (see Appendix \ref{sec:SOC_pert}):
+symmetric manifold $\M_\PS$ (see Appendix \ref{sec:SOC_pert}):
 \begin{align}
@@
 can be made homogenous through the gauge transformation
-$c_{j\mu}^\dag \to e^{i\mu\phi j} c_{j\mu}^\dag$, which takes
+$c_{j\mu}^\dag \to e^{-i\mu\phi j} c_{j\mu}^\dag$, which takes
 \begin{align}
@@
 $\Omega_{2,1}$ can be simultaneously tuned to zero by setting
-$\Omega_0=\Omega_+=\Omega_-$, at which point the only remaining
-effective drives in $H_{\t{Raman}}$ are $\sim\D_{1,1}$ and
-$\sim\D_{2,2}$ (as $\D_{0,0}\propto\1$ is just constant shift in
-energy).  If instead $\Omega_0=\pm\Omega_\pm$, then the amplitudes
-$\Omega_{1,0}$, $\Omega_{2,0}$ and $\Omega_{1,1}$ vanish, leaving only
-the drives $\D_{2,1}$ and $\D_{2,2}$.
+$\Omega_0=\Omega_+=\Omega_-=\xi_1\Omega$, at which point the Raman
+drive becomes (up to an overall constant)
+\begin{align}
+  H_{\t{Raman}}^{(+)} \equiv \f{\Omega^2}{\Delta}
+  \p{\sqrt{2}\, \col{S_\x} + \col{S_\x^2 - S_\y^2}},
+\end{align}
+where the single-spin operators $S_\alpha$ are related to the drive
+operators $D_{LM}$ through the relations in \eqref{eq:spin_ops},
+\eqref{eq:spin_ops_2_0}, and \eqref{eq:spin_ops_2_12}; and $\col{X}$
+is the collective version of $X$.  If instead
+$\Omega_0=\pm\Omega_\pm=\xi_1\Omega$, then up to a scalar term the
+drive becomes
+\begin{align}
+  H_{\t{Raman}}^{(-)} \equiv \f{\Omega^2}{\Delta}
+  \p{\sqrt{2}\, \sp{\col{S_\z S_\x+ S_\x S_\z}}
+    - \sp{\col{S_\x^2 - S_\y^2}}}.
+\end{align}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Weak SOC in the fully symmetric manifold}
+\section{Weak SOC in the PS manifold}
 \label{sec:SOC_pert}
@@
 Here we derive the first- and second-order effective Hamiltonians
-induced on the fully symmetric manifold $\M_\FS$ by the
-single-particle weak-SOC Hamiltonian
+induced on the PS manifold $\M_\PS$ by the single-particle weak-SOC
+Hamiltonian
 \begin{align}
@@
 % TWA dynamics
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: t
+%%% End:

### 2020-07-22 collective_sun/multilevel_spin_notes.tex
minor changes in some conventions

@@
     \sqrt{2} & M = 0 \\
-    \p{-1}^M \sign\p{M}^{M+1/2} & M \ne 0
+    \p{-1}^M & M > 0 \\
+    i & M < 0
   \end{cases},
@@
 \end{align}
-In particular, the amplitudes $\Omega_{1,0}$, $\Omega_{2,0}$, and
-$\Omega_{2,1}$ can be simultaneously tuned to zero by setting
-$\Omega_0=\Omega_+=\Omega_-=\xi_1\Omega$, at which point the Raman
-drive becomes (up to an overall constant)
+Note that $\D_{0,0}$ is proportional to the identity operator, so the
+amplitude $\Omega_{0,0}$ can be neglected without consequence.  If
+$\Omega_0=\Omega_+=\Omega_-=\xi_1\Omega$, then all remaining
+amplitudes except $\Omega_{1,1}$ and $\Omega_{2,2}$ vanish, and the
+Raman drive becomes
 \begin{align}
@@
 is the collective version of $X$.  If instead
-$\Omega_0=\pm\Omega_\pm=\xi_1\Omega$, then up to a scalar term the
-drive becomes
+$\Omega_0=\pm\Omega_\pm=\xi_1\Omega$, then all amplitudes except
+$\Omega_{2,1}$ and $\Omega_{2,2}$ vanish, and the drive becomes
 \begin{align}
@@
 \begin{align}
-  H_{\t{pd}}\p{t} = \sum_p \chi_p \omega \cos\p{\omega t} D_\z^{(p)},
+  H_{\t{pd}}\p{t} = \sum_p \chi_p \omega \cos\p{\omega t} S_\z^{(p)},
 \end{align}
@@
 driving amplitude for spin $p$, and
-$D_\z^{(p)} \equiv \xi_1 D_{1,0}^{(p)}$ with $\xi_1$ defined in
+$S_\z^{(p)} \equiv \xi_1 D_{1,0}^{(p)}$ with $\xi_1$ defined in
 \eqref{eq:scale_fac}.  The periodic drive $H_{\t{pd}}\p{t}$ can be
@@
 Transition operators $T_{LM}$ transform under rotations generated by
-the drive operator $D_\z$ as
-\begin{align}
-  e^{i\phi D_\z} T_{LM} e^{-i\phi D_\z} = e^{iM\phi} T_{LM}.
+the drive operator $S_\z$ as
+\begin{align}
+  e^{i\phi S_\z} T_{LM} e^{-i\phi S_\z} = e^{iM\phi} T_{LM}.
 \end{align}
@@
   U_{\t{pd}}\p{t} \equiv \exp\sp{-i\int_0^td\tau\,H_{\t{pd}}\p{\tau}}
-  = \prod_p \exp\sp{-i \chi_p \sin\p{\omega t} D_\z^{(p)}}.
+  = \prod_p \exp\sp{-i \chi_p \sin\p{\omega t} S_\z^{(p)}}.
 \end{align}

### 2020-07-22 collective_sun/sun_perturbation.tex
changes to figure names

@@
 processes through order $2\floor{r_{\t{max}}/M}+1$ in perturbation
-theory, when $\norm{\O\p{\m w,X}}<\Delta_{\t{gap}}/2$.  When
-$\norm{\O\p{\m w,X}}\ge\Delta_{\t{gap}}/2$, perturbation theory
-formally breaks down, and the error incurred by neglecting all
-$s$-magnon states with $s>r_{\t{max}}$ is not clear a priori.
-Nonetheless, TS$_{r_{\t{max}}}$ simulations can be benchmarked by
-examining the populations of all $r$-magnon manifolds $\M_r$ with
-$r\le r_{\t{max}}$.  As long as the maximal populations within
-large-$r$ manifolds $\M_r$ are small, one should expect
+theory.  When $\norm{\O\p{\m w,X}}\ge\Delta_{\t{gap}}/2$, however,
+perturbation theory formally breaks down, and the error incurred by
+neglecting all $s$-magnon states with $s>r_{\t{max}}$ is not clear a
+priori.  Nonetheless, TS$_{r_{\t{max}}}$ simulations can be
+benchmarked by examining the populations of all $r$-magnon manifolds
+$\M_r$ with $r\le r_{\t{max}}$.  As long as the maximal populations
+within large-$r$ manifolds $\M_r$ are small, one should expect
 TS$_{r_{\t{max}}}$ simulations to faithfully capture (at least
@@
   = \sum_{i,j} v_i w_j X_i Y_j
-  \EQPS \diagram{single_body_0} X_1 Y_2
-  + \diagram{single_body_1} X_1 Y_1,
+  \EQPS \diagram{prod_b11_c0} X_1 Y_2
+  + \diagram{prod_b11_c1} X_1 Y_1,
 \end{align}
@@
 \begin{align}
-  \diagram{single_body_0}
-  = \diagram{single_body_0_o} - \diagram{single_body_0_x}
-  = \diagram{single_body_0_oo} - \diagram{single_body_1}
+  \diagram{prod_b11_c0}
+  = \diagram{prod_b11_c0_o} - \diagram{prod_b11_c0_x}
+  = \diagram{prod_b11_c0_oo} - \diagram{prod_b11_c1}
   = \col{v}\,\col{w} - \m v \c \m w,
@@
   \O\p{\m v,X} \O\p{\m w,Y}
-  \EQPS \diagram{two_body_0} \spp{X,Y}_4 + \diagram{two_body_1}
-  \spp{X,Y}_3 + \diagram{two_body_2} \spp{X,Y}_2,
+  \EQPS \diagram{prod_b22_c0} \spp{X,Y}_4 + \diagram{prod_b22_c1}
+  \spp{X,Y}_3 + \diagram{prod_b22_c2} \spp{X,Y}_2,
   \label{eq:two_prod_pair}
@@
 \begin{align}
-  \diagram{two_body_1}
-  &= \sp{\diagram{two_body_1_o}} - \sp{\diagram{two_body_1_x}} \\
-  &= \sp{\diagram{two_body_1_oo} - \diagram{two_body_1_ox}}
-  - \sp{\diagram{two_body_2}} \label{eq:mid_zero} \\
-  &= \diagram{two_body_1_ooo} - \diagram{two_body_2_oo}
+  \diagram{prod_b22_c1}
+  &= \sp{\diagram{prod_b22_c1_o}} - \sp{\diagram{prod_b22_c1_x}} \\
+  &= \sp{\diagram{prod_b22_c1_oo} - \diagram{prod_b22_c1_ox}}
+  - \sp{\diagram{prod_b22_c2}} \label{eq:mid_zero} \\
+  &= \diagram{prod_b22_c1_ooo} - \diagram{prod_b22_c2_oo}
 \end{align}
@@
 \begin{align}
-  \diagram{two_body_0}
-  &= \sp{\diagram{two_body_0_o}} - \sp{\diagram{two_body_0_x}} \\
-  &= \sp{\diagram{two_body_0_oo} - \diagram{two_body_0_ox}}
-  - \sp{2 \diagram{two_body_1}} \\
-  &= \sp{\diagram{two_body_0_oooo} - 2 \diagram{two_body_1_o}}
-  - \sp{2 \diagram{two_body_1_ooo} - 2 \diagram{two_body_2_oo}} \\
-  &= \diagram{two_body_0_oooo} - 4 \diagram{two_body_1_ooo}
-  + 2 \diagram{two_body_2_oo}.
+  \diagram{prod_b22_c0}
+  &= \sp{\diagram{prod_b22_c0_o}} - \sp{\diagram{prod_b22_c0_x}} \\
+  &= \sp{\diagram{prod_b22_c0_oo} - \diagram{prod_b22_c0_ox}}
+  - \sp{2 \diagram{prod_b22_c1}} \\
+  &= \sp{\diagram{prod_b22_c0_oooo} - 2 \diagram{prod_b22_c1_o}}
+  - \sp{2 \diagram{prod_b22_c1_ooo} - 2 \diagram{prod_b22_c2_oo}} \\
+  &= \diagram{prod_b22_c0_oooo} - 4 \diagram{prod_b22_c1_ooo}
+  + 2 \diagram{prod_b22_c2_oo}.
 \end{align}
@@
 \begin{align}
-  \diagram{two_body_2} = \v{\m v} \c\v{\m w},
-  &&
-  \diagram{two_body_1} =
+  \diagram{prod_b22_c2} = \v{\m v} \c\v{\m w},
+  &&
+  \diagram{prod_b22_c1} =
   \v v \c \v w - \v{\m v} \c \v{\m w},
   &&
-  \diagram{two_body_0}
+  \diagram{prod_b22_c0}
   = \col{v}\,\col{w} - 4 \v v\c\v w + 2 \v{\m v}\c\v{\m w}.
@@
 \begin{align}
-  A_4 \equiv \diagram{two_body_0}
+  A_4 \equiv \diagram{prod_b22_c0}
   = \col{v}\,\col{w} - 4 \v v\c\v w + 2 \v{\m v}\c\v{\m w},
   &&
-  A_2 \equiv 4 \diagram{two_body_1}
+  A_2 \equiv 4 \diagram{prod_b22_c1}
   = 4 \v v \c \v w - 4 \v{\m v} \c \v{\m w}
@@
 \begin{align}
-  A_0 \equiv 2 \diagram{two_body_2} = 2 \v{\m v} \c\v{\m w}.
+  A_0 \equiv 2 \diagram{prod_b22_c2} = 2 \v{\m v} \c\v{\m w}.
 \end{align}
@@
 is a falling factorial.
+
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \subsection{Three single-body operators}
+% \label{sec:single_trip_prod}
+
+% Here we simplify the single-body product
+% \begin{align}
+%   \O\p{\m u,X} \O\p{\m v,Y} \O\p{\m w,Z}
+%   = \sum_{i,j,k} u_i v_j w_k X_i Y_j Z_k,
+% \end{align}
+% where $i,j,k$ index individual spins that the single-spin operators
+% $X,Y,K$ act on; and $\m m\equiv\sum_i m_i\ket{i}$ for
+% $\m m\in\set{\m u,\m v,\m w}$ are vectors of scalar coefficients.
+% This product appears, for example, in the calculation of matrix
+% elements of the operator $\O\p{\m v,Y}$ within the single-magnon
+% manifold.  The diagrammatic expansion of this operator is
+% \begin{multline}
+%   \O\p{\m u,X} \O\p{\m v,Y} \O\p{\m w,Z}
+%   \EQPS \diagram{prod_b111_c0} X_1 Y_2 Z_3
+%   + \diagram{prod_b111_c10} X_1 Y_1 Z_1 \\
+%   + \diagram{prod_b111_c11_u} X_1 Y_2 Z_2
+%   + \diagram{prod_b111_c11_v} X_2 Y_1 Z_2
+%   + \diagram{prod_b111_c11_w} X_2 Y_2 Z_1,
+% \end{multline}
+% where we can reduce (by hand or with a computer)
 
@@
 \begin{align}
-  A_6 \equiv \diagram{triple_0},
-  &&
-  A_4 \equiv 4 \diagram{triple_01} + 8 \diagram{triple_1},
-  &&
-  A_0 \equiv 8 \diagram{triple_0111},
+  A_6 \equiv \diagram{prod_b222_c0},
+  &&
+  A_4 \equiv 4 \diagram{prod_b222_c01} + 8 \diagram{prod_b222_c1},
+  &&
+  A_0 \equiv 8 \diagram{prod_b222_c0111},
   \label{eq:triple_64}
@@
 \begin{align}
-  A_2 \equiv 8 \diagram{triple_011} + 2 \diagram{triple_02}
-  + 8 \diagram{triple_11} + 4 \diagram{triple_2},
+  A_2 \equiv 8 \diagram{prod_b222_c011} + 2 \diagram{prod_b222_c02}
+  + 8 \diagram{prod_b222_c11} + 4 \diagram{prod_b222_c2},
   \label{eq:triple_2}
@@
 \begin{align}
-  \diagram{triple_1} \equiv \diagram{triple_1_uvw},
+  \diagram{prod_b222_c1} \equiv \diagram{prod_b222_c1_uvw},
   \label{eq:triple_uvw}
@@
 \begin{align}
-  \diagram{triple_011}
-  \equiv \diagram{triple_011_uvw}
-  + \diagram{triple_011_vwu} + \diagram{triple_011_wuv}.
+  \diagram{prod_b222_c011}
+  \equiv \diagram{prod_b222_c011_uvw}
+  + \diagram{prod_b222_c011_vwu} + \diagram{prod_b222_c011_wuv}.
   \label{eq:triple_uvw_3}
@@
 \begin{align}
-  D_6 \equiv \diagram{triple_0_o},
-  &&
-  D_4 \equiv 4 \diagram{triple_01_o} - 16 \diagram{triple_1_o},
-  &&
-  D_0 \equiv 8 \diagram{triple_0111_o},
-\end{align}
-\begin{align}
-  D_2 \equiv 8 \diagram{triple_011_o}
-  + 2 \diagram{triple_02_o}
-  - 16 \diagram{triple_11_o}
-  + 16 \diagram{triple_2_o}.
+  D_6 \equiv \diagram{prod_b222_c0_o},
+  &&
+  D_4 \equiv 4 \diagram{prod_b222_c01_o} - 16 \diagram{prod_b222_c1_o},
+  &&
+  D_0 \equiv 8 \diagram{prod_b222_c0111_o},
+\end{align}
+\begin{align}
+  D_2 \equiv 8 \diagram{prod_b222_c011_o}
+  + 2 \diagram{prod_b222_c02_o}
+  - 16 \diagram{prod_b222_c11_o}
+  + 16 \diagram{prod_b222_c2_o}.
 \end{align}

### 2020-07-22 collective_sun/scripts/simplify_product_diagrams.py

@@ -7,7 +7,7 @@ import copy
 from itertools_extension import set_diagrams
 from operator_product_methods import diagram_vec
 
-dimensions = [ 2, 2 ]
+dimensions = [ 2, 2, 2 ]
 
 def num_assignments(diagram):
     factors = set.union(*[ set(subset) for subset in diagram.keys() ])

### 2020-07-30 sun_phases/sun_dynamics_notes.tex
added calculation of S^2 for the single-particle Hamiltonian

\documentclass[nofootinbib,notitlepage,11pt]{revtex4-2}

%%% linking references
\usepackage{hyperref}
\hypersetup{
  breaklinks=true,
  colorlinks=true,
  linkcolor=blue,
  filecolor=magenta,
  urlcolor=cyan,
}

%%% header / footer
\usepackage{fancyhdr} % easier header and footer management
\pagestyle{fancy} % page formatting style
\fancyhf{} % clear all header and footer text
\renewcommand{\headrulewidth}{0pt} % remove horizontal line in header
\usepackage{lastpage} % for referencing last page
\cfoot{\thepage~of \pageref{LastPage}} % "x of y" page labeling

%%% symbols, notations, etc.
\usepackage{physics,braket,bm,amssymb} % physics and math
\renewcommand{\t}{\text} % text in math mode
\newcommand{\f}[2]{\dfrac{#1}{#2}} % shorthand for fractions
\newcommand{\p}[1]{\left(#1\right)} % parenthesis
\renewcommand{\sp}[1]{\left[#1\right]} % square parenthesis
\renewcommand{\set}[1]{\left\{#1\right\}} % curly parenthesis
\newcommand{\bk}{\Braket} % shorthand for braket notation
\renewcommand{\v}{\bm} % bold vectors
\newcommand{\uv}[1]{\bm{\hat{#1}}} % unit vectors
\newcommand{\av}{\vec} % arrow vectors
\renewcommand{\c}{\cdot} % inner product
\renewcommand{\d}{\partial} % partial derivative
\renewcommand{\dd}{\text{d}} % for infinitesimals
\renewcommand{\i}{\mathrm{i}\mkern1mu} % imaginary unit

\usepackage{dsfont} % for identity operator
\newcommand{\1}{\mathds{1}}

\usepackage{mathtools} % for \coloneqq

\newcommand{\up}{\uparrow}
\newcommand{\dn}{\downarrow}

\newcommand{\x}{\text{x}}
\newcommand{\y}{\text{y}}
\newcommand{\z}{\text{z}}

\newcommand{\B}{\mathcal{B}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathcal{E}}
\renewcommand{\H}{\mathcal{H}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\J}{\mathcal{J}}
\newcommand{\M}{\mathcal{M}}
\newcommand{\N}{\mathcal{N}}
\renewcommand{\O}{\mathcal{O}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\T}{\mathcal{T}}
\renewcommand{\S}{\mathcal{S}}
\newcommand{\V}{\mathcal{V}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\Z}{\mathcal{Z}}

\newcommand{\EE}{\mathbb{E}}
\renewcommand{\SS}{\mathbb{S}}
\newcommand{\ZZ}{\mathbb{Z}}

\newcommand{\PS}{\text{PS}}
\newcommand{\col}{\underline}

\usepackage[inline]{enumitem} % in-line lists and \setlist{} (below)
\setlist[enumerate,1]{label={(\roman*)}} % default in-line numbering
\setlist{nolistsep} % more compact spacing between environments

%%% text markup
\usepackage{color} % text color
\newcommand{\red}[1]{{\color{red} #1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\thispagestyle{fancy}

% todo: find roots of lax vector

\title{Collective SU($n$) spin model dynamics}%
\author{Michael A. Perlin}%
\date{\today}

\maketitle

We first consider a general quadratic spin Hamiltonian of the form
\begin{align}
  H_{\t{spin}} = \f1N \sum_{\substack{\mu,\nu,\rho,\sigma\\i<j}}
  h^{\mu\nu i}_{\rho\sigma j} s_{\mu\nu i} s_{\rho\sigma j}
  + \sum_{\mu,\nu,i} \epsilon_{\mu\nu i} s_{\mu\nu i},
  \label{eq:spin}
\end{align}
where $\mu,\nu\in\ZZ_n$ index orthogonal states of a single spin;
$i,j\in\ZZ_N$ index one of $N$ spins; $h^{\mu\nu i}_{\rho\sigma j}$
and $\epsilon_{\mu\nu i}$ are scalars; and
$s_{\mu\nu i}\coloneqq\op{\mu}{\nu}_i$ is a transition operator for
spin $i$.  If interactions are SU($n$) symmetric then
$h^{\mu\nu i}_{\rho\sigma k} = h_{ik}
\delta_{\mu\sigma}\delta_{\nu\rho}$, in which case
\begin{align}
  H_{\t{spin}} = \f1N \sum_{\substack{\mu,\nu,\rho,\sigma\\i<j}}
  h_{ij} \Pi_{ij}
  + \sum_{\mu,\nu,i} \epsilon_{\mu\nu i} s_{\mu\nu i},
  &&
  \Pi_{ij} \equiv \sum_{\mu,\nu} s_{\mu\nu i} s_{\rho\sigma j}.
  \label{eq:spin_sun}
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\section{Schwinger boson representation}

We can write the Hamiltonian in \eqref{eq:spin} using Schwinger bosons
as
\begin{align}
  H_{\t{boson}} = \f1N \sum_{\substack{\mu,\nu,\rho,\sigma\\i<j}}
  h^{\mu\nu i}_{\rho\sigma j}
  b_{\mu i}^\dag b_{\nu i} b_{\rho j}^\dag b_{\sigma j}
  + \sum_{\mu,\nu,i} \epsilon_{\mu\nu i} b_{\mu i}^\dag b_{\nu i},
\end{align}
where $b_{\mu i}$ annihilates a boson of type $\mu$ on site $i$.  The
Heisenberg equations of motion for these operators are\footnote{The
  Hamiltonian in \eqref{eq:spin} only defines
  $h^{\mu\nu j}_{\rho\sigma k}$ for $j<k$.  To simplify expressions in
  the remainder of this work, we therefore additionally define
  $h^{\mu\nu k}_{\rho\sigma j}=h^{\rho\sigma j}_{\mu\nu k}$ and
  $h^{\mu\nu i}_{\rho\sigma i}=0$.}
\begin{align}
  \i \d_t b_{\alpha i} = \sp{b_{\alpha i}, H_{\t{boson}}}
  &= \f1N \sum_{\substack{\mu,\nu,\rho,\sigma\\j<k}}
  h^{\mu\nu j}_{\rho\sigma k}
  \sp{b_{\alpha i}, b_{\mu j}^\dag b_{\nu j} b_{\rho k}^\dag b_{\sigma k}}
  + \sum_{\mu,\nu,j} \epsilon_{\mu\nu j}
  \sp{b_{\alpha i}, b_{\mu j}^\dag b_{\nu j}} \\
  &= \f1N \sum_{\mu,\nu,\rho,\sigma,k}
  h^{\mu\nu j}_{\rho\sigma k}
  \sp{b_{\alpha i}, b_{\mu i}^\dag b_{\nu i}} b_{\rho k}^\dag b_{\sigma k}
  + \sum_{\mu,\nu} \epsilon_{\mu\nu i}
  \sp{b_{\alpha i}, b_{\mu i}^\dag b_{\nu i}} \\
  &= \sum_{\mu,\nu} \p{\f1N \sum_{\rho,\sigma,k}
    h^{\mu\nu i}_{\rho\sigma k} b_{\rho k}^\dag b_{\sigma k}
    + \epsilon_{\mu\nu i}}
  \sp{b_{\alpha i}, b_{\mu i}^\dag b_{\nu i}}
\end{align}
where
\begin{align}
  \sp{b_{\alpha i}, b_{\mu i}^\dag b_{\nu i}}
  = \delta_{\mu\alpha} \delta_{\nu\alpha} b_{\alpha i}
  + \delta_{\mu\alpha} \p{1-\delta_{\nu\alpha}} b_{\nu i}
  = \delta_{\mu\alpha} b_{\nu i},
\end{align}
so
\begin{align}
  \i \d_t b_{\alpha i}
  = \sum_\nu \p{\f1N \sum_{\rho,\sigma,k}
    h^{\alpha\nu i}_{\rho\sigma k} b_{\rho k}^\dag b_{\sigma k}
    + \epsilon_{\alpha\nu i}} b_{\nu i}.
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Non-interacting limit of a realistic 1D model}

Consider the single-particle Hamiltonian
\begin{align}
  H = -\f{J}{2} \sum_{\mu,q} \cos\p{q + \mu\phi} s_{\mu\mu q},
\end{align}
where $q$ is a quasi-momentum (in units with lattice spacing $a=1$);
$\mu\in\set{-I,-I+1,\cdots,I}$ is a nuclear spin projection onto a
quantization axis, with $I\equiv\p{n-1}/2$ the total nuclear spin;
$\phi$ is a spin-orbit-coupling angle, and $J$ is the single-particle
bandwidth (proportional to the tunneling rate, $J/4$).  We wish to
compute an expectation value of the operator
\begin{align}
  S^2\p{t}
  \equiv e^{it H} \p{\v S^\dag \c\v S} e^{-it H}
  = \sum_{\mu,\nu,p,q} s_{\mu\nu p} s_{\nu\mu q} \times
  e^{itJ\p{\eta_{\mu\nu p}-\eta_{\mu\nu q}}},
\end{align}
where
\begin{align}
  \eta_{\mu\nu p}
  \equiv -\f12 \sp{\cos\p{p+\mu\phi} - \cos\p{p+\nu\phi}}
  = \sin\p{\phi_{\mu\nu}^-} \sin\p{p+\phi_{\mu\nu}^+},
  &&
  \phi_{\mu\nu}^\pm \equiv \f{\mu\pm\nu}{2}\,\phi.
\end{align}
If $\ket\Psi=\ket\psi^{\otimes N}$, then
\begin{align}
  \bk{\Psi|S^2\p{t}|\Psi}
  &= \sum_{\mu,\nu} \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2
  \sum_{p\ne q} e^{itJ\p{\eta_{\mu\nu p}-\eta_{\mu\nu q}}} + n N \\
  &= \sum_{\mu,\nu} \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2
  \sum_{p,q} e^{itJ\p{\eta_{\mu\nu p}-\eta_{\mu\nu q}}} + \p{n-1} N \\
  &= N^2 \sum_{\mu,\nu} \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2
  \abs{\xi_{\mu\nu}\p{\phi}}^2 + \p{n-1} N,
\end{align}
where
\begin{align}
  \xi_{\mu\nu}\p{\phi} \equiv \f1N \sum_q e^{itJ\eta_{\mu\nu q}}
  = \f1N \sum_q e^{itJ\sin\p{\phi_{\mu\nu}^-} \sin\p{q+\phi_{\mu\nu}^+}}.
\end{align}
As $N\to\infty$, this sum is well approximated by an integral:
\begin{align}
  \xi_{\mu\nu}\p{\phi} \approx \f1{2\pi} \int_{-\pi}^\pi \dd\theta\,
  e^{itJ\sin\p{\phi_{\mu\nu}^-} \sin\p{\theta+\phi_{\mu\nu}^+}}
  = \J_0\sp{tJ\sin\p{\phi_{\mu\nu}^-}},
\end{align}
where $\J_0$ is the zero-order Bessel function of the first kind.  To
leading order in $N\gg1$, we thus find that
\begin{align}
  \bk{\Psi|S^2\p{t}|\Psi}
  \approx N^2 \sum_{\mu,\nu} \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2
  \J_0\sp{tJ\sin\p{\phi_{\mu\nu}^-}}^2.
\end{align}
The time-independent terms in the sum have $\mu=\nu$ (in which case
$\phi_{\mu\nu}^-=0$), so
\begin{align}
  \bk{\Psi|S^2_{\t{const}}|\Psi}
  \approx N^2 \sum_{\mu,\mu} \abs{\bk{\psi\op{\mu}{\mu}\psi}}^2
  = N^2 r,
  &&
  r \equiv \sum_\mu \abs{\bk{\mu|\psi}}^4.
  \label{eq:S^2_const}
\end{align}
Up to an inconsequential choice of gauge, we can prepare a state of
the form
\begin{align}
  \ket\psi = \f1{\sqrt{2}}
  \p{e^{i\beta} \ket{\theta,\alpha}
    + e^{-i\beta} \ket{\theta,-\alpha}},
  &&
  \ket{\theta,\alpha} \equiv e^{i\alpha s_\z} e^{i\theta s_\y} \ket\up,
\end{align}
where $\theta,\alpha$ are respectively polar and azimuthal angles of
the polarized state $\ket{\theta,\alpha}$ of a single spin; the spin
operators $s_\z,s_\y$ generate rotations about the $z,y$ axes, with
e.g.~$s_\z\equiv\sum_\mu \mu\op\mu$; and $\ket\up$ is a state
polarized along the $z$ axis.  More specifically, the state that we
can prepare has
\begin{align}
  \theta = \f{\pi}{2} + \arcsin\p{\f13},
  &&
  \alpha = \f{\pi}{3},
  &&
  \beta = -\f{2\pi}{3}\, I.
\end{align}
In order to evaluate \eqref{eq:S^2_const}, we therefore
expand\footnote{The expansion in \eqref{eq:polarized_proj} is acquired
  by representing an $n$-level spin by $n-1$ qubits restricted to the
  (Dicke) manifold of permutationally symmetric states.}
\begin{align}
  \bk{\mu|\theta,\alpha} = { 2I \choose I+\mu }^{1/2}
  \cos\p{\f{\theta}{2}}^{I+\mu} \sin\p{\f{\theta}{2}}^{I-\mu}
  e^{-i\mu\alpha},
  \label{eq:polarized_proj}
\end{align}
which implies that
\begin{align}
  r = \sum_\mu { 2I \choose I+\mu }^2
  \sp{\cos\p{\f{\theta}{2}}^{I+\mu} \sin\p{\f{\theta}{2}}^{I-\mu}}^4
  \times 4 \cos\p{\mu\alpha-\beta}^4.
\end{align}
Figure \ref{fig:const_vals} shows several values of $r$ for different
$n$.

\begin{figure}
  \centering
  \includegraphics{figures/oscillations/const_vals.pdf}
  \caption{Several values of
    $r \equiv\sum_\mu\abs{\bk{\mu|\psi}}^4 =
    \lim_{N\to\infty}\bk{\Psi|S^2_{\t{const}}|\Psi}/N^2$ for different
    $n$.}
  \label{fig:const_vals}
\end{figure}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2020-07-30 sun_phases/sun_dynamics_notes.tex
changed conventions for order parameter and add a small-angle approximation

@@
 \newcommand{\col}{\underline}
+
+\let\var\relax
+\DeclareMathOperator{\var}{var}
 
@@
 \begin{align}
-  S^2\p{t}
-  \equiv e^{it H} \p{\v S^\dag \c\v S} e^{-it H}
-  = \sum_{\mu,\nu,p,q} s_{\mu\nu p} s_{\nu\mu q} \times
+  \O\p{t}
+  \equiv e^{it H} \p{\f{\v S^\dag \c\v S}{N^2}} e^{-it H}
+  = \f1{N^2} \sum_{\mu,\nu,p,q} s_{\mu\nu p} s_{\nu\mu q} \times
   e^{itJ\p{\eta_{\mu\nu p}-\eta_{\mu\nu q}}},
@@
 \begin{align}
-  \bk{\Psi|S^2\p{t}|\Psi}
+  \bk{\Psi|\O\p{t}|\Psi}
+  &= \f1{N^2} \sum_{\mu,\nu} \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2
+  \sum_{p\ne q} e^{itJ\p{\eta_{\mu\nu p}-\eta_{\mu\nu q}}} + \f{n}{N} \\
+  &= \f1{N^2} \sum_{\mu,\nu} \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2
+  \sum_{p,q} e^{itJ\p{\eta_{\mu\nu p}-\eta_{\mu\nu q}}} + \f{n-1}{N} \\
   &= \sum_{\mu,\nu} \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2
-  \sum_{p\ne q} e^{itJ\p{\eta_{\mu\nu p}-\eta_{\mu\nu q}}} + n N \\
-  &= \sum_{\mu,\nu} \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2
-  \sum_{p,q} e^{itJ\p{\eta_{\mu\nu p}-\eta_{\mu\nu q}}} + \p{n-1} N \\
-  &= N^2 \sum_{\mu,\nu} \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2
-  \abs{\xi_{\mu\nu}\p{\phi}}^2 + \p{n-1} N,
+  \abs{\xi_{\mu\nu}\p{\phi}}^2 + \f{n-1}{N},
 \end{align}
@@
 \begin{align}
-  \bk{\Psi|S^2\p{t}|\Psi}
-  \approx N^2 \sum_{\mu,\nu} \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2
+  \bk{\Psi|\O\p{t}|\Psi}
+  \approx \sum_{\mu,\nu} \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2
   \J_0\sp{tJ\sin\p{\phi_{\mu\nu}^-}}^2.
+  \label{eq:O_time}
 \end{align}
+
+\subsection{Long-time average}
+
 The time-independent terms in the sum have $\mu=\nu$ (in which case
@@
 \begin{align}
-  \bk{\Psi|S^2_{\t{const}}|\Psi}
-  \approx N^2 \sum_{\mu,\mu} \abs{\bk{\psi\op{\mu}{\mu}\psi}}^2
-  = N^2 r,
-  &&
-  r \equiv \sum_\mu \abs{\bk{\mu|\psi}}^4.
-  \label{eq:S^2_const}
+  \bk{\Psi|\O_{\t{const}}|\Psi}
+  \approx r \equiv \sum_\mu \abs{\bk{\mu|\psi}}^4.
+  \label{eq:O_const}
 \end{align}
-Up to an inconsequential choice of gauge, we can prepare a state of
-the form
+where $r$ is sometimes called the inverse participation ratio of
+$\ket\psi$ with respect to the basis $\set{\ket\mu}$.  Up to an
+inconsequential choice of gauge, we can prepare a state of the form
 \begin{align}
@@
 \end{align}
-In order to evaluate \eqref{eq:S^2_const}, we therefore
+In order to evaluate \eqref{eq:O_const}, we therefore
 expand\footnote{The expansion in \eqref{eq:polarized_proj} is acquired
@@
 \begin{align}
-  \bk{\mu|\theta,\alpha} = { 2I \choose I+\mu }^{1/2}
-  \cos\p{\f{\theta}{2}}^{I+\mu} \sin\p{\f{\theta}{2}}^{I-\mu}
-  e^{-i\mu\alpha},
+  \bk{\mu|\theta,\alpha} = \chi_\mu e^{-i\mu\alpha},
+  &&
+  \chi_\mu \equiv { 2I \choose I+\mu }^{1/2}
+  \cos\p{\f{\theta}{2}}^{I+\mu} \sin\p{\f{\theta}{2}}^{I-\mu},
   \label{eq:polarized_proj}
@@
 \begin{align}
-  r = \sum_\mu { 2I \choose I+\mu }^2
-  \sp{\cos\p{\f{\theta}{2}}^{I+\mu} \sin\p{\f{\theta}{2}}^{I-\mu}}^4
-  \times 4 \cos\p{\mu\alpha-\beta}^4.
+  \bk{\mu|\psi} = \chi_\mu \times \sqrt{2}\, \cos\p{\mu\alpha-\beta},
+  &&
+  r = \sum_\mu \chi_\mu^4 \times 4 \cos\p{\mu\alpha-\beta}^4.
 \end{align}
-Figure \ref{fig:const_vals} shows several values of $r$ for different
-$n$.
+Figure \ref{fig:limiting_vals} shows several values of $r$ for different $n$.
 
@@
   \includegraphics{figures/oscillations/const_vals.pdf}
-  \caption{Several values of
-    $r \equiv\sum_\mu\abs{\bk{\mu|\psi}}^4 =
-    \lim_{N\to\infty}\bk{\Psi|S^2_{\t{const}}|\Psi}/N^2$ for different
+  \includegraphics{figures/oscillations/var_z_vals.pdf}
+  \caption{Several values of $r$ and $\var\p{s_\z/I}$ for different
     $n$.}
-  \label{fig:const_vals}
+  \label{fig:limiting_vals}
 \end{figure}
+
+\subsection{Small-angle limit}
+
+In the limit of a small spin-orbit coupling angle, $\phi I\ll1$, as
+long as $tJ\ll\p{\phi I}^{-1}$ we can expand the Bessel function in
+\eqref{eq:O_time} as $\J_0\p{\tau}^2\approx1-\tau^2/2+O\p{\tau^4}$ and
+simplify the resulting expression to get
+\begin{align}
+  \bk{\Psi|\O\p{t}|\Psi}
+  \approx 1 - \f14 \var\p{s_\z} \p{tJ\phi}^2,
+  &&
+  \var\p{X} \equiv \bk{\psi|X^2|\psi} - \bk{\psi|X|\psi}^2.
+\end{align}
+Figure \ref{fig:limiting_vals} shows several values of
+$\var\p{s_\z/I}$ for different $n$.
 

### 2020-07-30 sun_phases/scripts/plot_limiting_vals.py
add script to make some plots in sun_dynamics_notes

@@ -1,0 +1,80 @@
+#!/usr/bin/env python3
+
+import numpy as np
+import matplotlib.pyplot as plt
+import scipy.special
+
+from dicke_methods import coherent_spin_state_angles
+
+figsize = (3,2)
+
+params = { "font.size" : 12,
+           "text.usetex" : True,
+           "text.latex.preamble" : [ r"\usepackage{physics}" ]}
+plt.rcParams.update(params)
+
+##############################
+
+n_vals = np.arange(2,21,2)
+S_vals = (n_vals-1)/2
+def spin_vals(n):
+    S = (n-1)/2
+    return np.linspace(-S,S,n)
+
+theta = np.pi/2 + np.arcsin(1/3)
+alpha = np.pi/3
+def beta(n):
+    return -(n-1)/3 * np.pi
+def state(n):
+    return ( coherent_spin_state_angles(theta, +alpha, n-1) * np.exp(+1j*beta(n)) +
+             coherent_spin_state_angles(theta, -alpha, n-1) * np.exp(-1j*beta(n)) ) / np.sqrt(2)
+
+##############################
+
+def chi(m,n):
+    S = (n-1)/2
+    return ( np.sqrt(scipy.special.binom(2*S,S+m)) *
+             np.cos(theta/2)**(S+m) * np.sin(theta/2)**(S-m) )
+def proj(m,n):
+    return chi(m,n) * np.sqrt(2) * np.cos(m*alpha-beta(n))
+
+for n in n_vals:
+    assert(np.allclose( state(n), [ proj(m,n) for m in spin_vals(n) ] ))
+
+##############################
+
+def r_val(n):
+    return sum( chi(m,n)**4 * 4 * np.cos(m*alpha-beta(n))**4
+                for m in spin_vals(n) )
+
+r_vals = [ sum(abs(state(n))**4) for n in n_vals ]
+r_vals_alt = [ r_val(n) for n in n_vals ]
+assert(np.allclose(r_vals, r_vals_alt))
+
+plt.figure(figsize = figsize)
+plt.plot(n_vals, r_vals, "k.")
+plt.gca().set_ylim(bottom = 0)
+plt.xlabel("$n$")
+plt.ylabel("$r$")
+plt.tight_layout()
+plt.savefig("../figures/oscillations/const_vals.pdf")
+plt.close("all")
+
+##############################
+
+def var_sz(n):
[…]

### 2020-07-30 sun_phases/sun_dynamics_notes.tex
add script to make some plots in sun_dynamics_notes
rearranged some notes

@@
 \begin{align}
-  H_{\t{spin}} = \f1N \sum_{\substack{\mu,\nu,\rho,\sigma\\i<j}}
+  H = \f1N \sum_{\substack{\mu,\nu,\rho,\sigma\\i<j}}
   h^{\mu\nu i}_{\rho\sigma j} s_{\mu\nu i} s_{\rho\sigma j}
@@
 \begin{align}
-  H_{\t{spin}} = \f1N \sum_{\substack{\mu,\nu,\rho,\sigma\\i<j}}
+  H = \f1N \sum_{\substack{\mu,\nu,\rho,\sigma\\i<j}}
   h_{ij} \Pi_{ij}
@@
 \begin{align}
-  H_{\t{boson}} = \f1N \sum_{\substack{\mu,\nu,\rho,\sigma\\i<j}}
+  H = \f1N \sum_{\substack{\mu,\nu,\rho,\sigma\\i<j}}
   h^{\mu\nu i}_{\rho\sigma j}
@@
 \begin{align}
-  \i \d_t b_{\alpha i} = \sp{b_{\alpha i}, H_{\t{boson}}}
+  \i \d_t b_{\alpha i} = \sp{b_{\alpha i}, H}
   &= \f1N \sum_{\substack{\mu,\nu,\rho,\sigma\\j<k}}
@@
   \O\p{t}
-  \equiv e^{it H} \p{\f{\v S^\dag \c\v S}{N^2}} e^{-it H}
+  \equiv e^{\i t H} \p{\f{\v S^\dag \c\v S}{N^2}} e^{-\i t H}
   = \f1{N^2} \sum_{\mu,\nu,p,q} s_{\mu\nu p} s_{\nu\mu q} \times
-  e^{itJ\p{\eta_{\mu\nu p}-\eta_{\mu\nu q}}},
+  e^{\i tJ\p{\eta_{\mu\nu p}-\eta_{\mu\nu q}}},
 \end{align}
@@
   &= \f1{N^2} \sum_{\mu,\nu} \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2
-  \sum_{p\ne q} e^{itJ\p{\eta_{\mu\nu p}-\eta_{\mu\nu q}}} + \f{n}{N} \\
-  &= \f1{N^2} \sum_{\mu,\nu} \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2
-  \sum_{p,q} e^{itJ\p{\eta_{\mu\nu p}-\eta_{\mu\nu q}}} + \f{n-1}{N} \\
+  \sum_{p\ne q} e^{\i tJ\p{\eta_{\mu\nu p}-\eta_{\mu\nu q}}} + \f{n}{N} \\
   &= \sum_{\mu,\nu} \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2
@@
 \begin{align}
-  \xi_{\mu\nu}\p{\phi} \equiv \f1N \sum_q e^{itJ\eta_{\mu\nu q}}
-  = \f1N \sum_q e^{itJ\sin\p{\phi_{\mu\nu}^-} \sin\p{q+\phi_{\mu\nu}^+}}.
+  \xi_{\mu\nu}\p{\phi} \equiv \f1N \sum_q e^{\i tJ\eta_{\mu\nu q}}
+  = \f1N \sum_q e^{\i tJ\sin\p{\phi_{\mu\nu}^-} \sin\p{q+\phi_{\mu\nu}^+}}.
 \end{align}
@@
   \xi_{\mu\nu}\p{\phi} \approx \f1{2\pi} \int_{-\pi}^\pi \dd\theta\,
-  e^{itJ\sin\p{\phi_{\mu\nu}^-} \sin\p{\theta+\phi_{\mu\nu}^+}}
+  e^{\i tJ\sin\p{\phi_{\mu\nu}^-} \sin\p{\theta+\phi_{\mu\nu}^+}}
   = \J_0\sp{tJ\sin\p{\phi_{\mu\nu}^-}},
@@
 
-\subsection{Long-time average}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Initial state}
 
-The time-independent terms in the sum have $\mu=\nu$ (in which case
-$\phi_{\mu\nu}^-=0$), so
-\begin{align}
-  \bk{\Psi|\O_{\t{const}}|\Psi}
-  \approx r \equiv \sum_\mu \abs{\bk{\mu|\psi}}^4.
-  \label{eq:O_const}
-\end{align}
-where $r$ is sometimes called the inverse participation ratio of
-$\ket\psi$ with respect to the basis $\set{\ket\mu}$.  Up to an
-inconsequential choice of gauge, we can prepare a state of the form
+Up to an inconsequential choice of gauge, the initial state that we
+can prepare has the form
 \begin{align}
   \ket\psi = \f1{\sqrt{2}}
-  \p{e^{i\beta} \ket{\theta,\alpha}
-    + e^{-i\beta} \ket{\theta,-\alpha}},
+  \p{e^{\i\beta} \ket{\theta,\alpha}
+    + e^{-\i\beta} \ket{\theta,-\alpha}},
   &&
   \ket{\theta,\alpha} \equiv e^{i\alpha s_\z} e^{i\theta s_\y} \ket\up,
+  \label{eq:double_state}
 \end{align}
@@
 \end{align}
-In order to evaluate \eqref{eq:O_const}, we therefore
+In order to compute expectation values with respect to $\ket\psi$, we
 expand\footnote{The expansion in \eqref{eq:polarized_proj} is acquired
@@
 \begin{align}
-  \bk{\mu|\theta,\alpha} = \chi_\mu e^{-i\mu\alpha},
+  \ket{\theta,\alpha} = \sum_\mu \chi_\mu e^{-\i\mu\alpha} \ket\mu,
   &&
@@
 \begin{align}
-  \bk{\mu|\psi} = \chi_\mu \times \sqrt{2}\, \cos\p{\mu\alpha-\beta},
-  &&
-  r = \sum_\mu \chi_\mu^4 \times 4 \cos\p{\mu\alpha-\beta}^4.
+  \ket\psi = \sum_\mu \chi_\mu \times \sqrt{2}\,
+  \cos\p{\mu\alpha-\beta} \ket\mu.
 \end{align}
-Figure \ref{fig:limiting_vals} shows several values of $r$ for different $n$.
 
-\begin{figure}
-  \centering
-  \includegraphics{figures/oscillations/const_vals.pdf}
-  \includegraphics{figures/oscillations/var_z_vals.pdf}
-  \caption{Several values of $r$ and $\var\p{s_\z/I}$ for different
-    $n$.}
-  \label{fig:limiting_vals}
-\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Limiting cases}
 
-\subsection{Small-angle limit}
+We now return to the order parameter $\O\p{t}$ in \eqref{eq:O_time}.
+The constant (time independent) terms in this expression have
+$\mu=\nu$ (in which case all $\phi_{\mu\nu}^-=0$), so up to $O(1/N)$
+corrections
+\begin{align}
+  \bk{\Psi|\O_{\t{const}}|\Psi}
+  \approx r \equiv \sum_\mu \abs{\bk{\mu|\psi}}^4,
+  \label{eq:O_const}
+\end{align}
+which is sometimes called the inverse participation ratio of
+$\ket\psi$ with respect to the basis $\set{\ket\mu}$.  Figure
+\ref{fig:limiting_vals} shows several values of $r$ for different $n$.
 
-In the limit of a small spin-orbit coupling angle, $\phi I\ll1$, as
-long as $tJ\ll\p{\phi I}^{-1}$ we can expand the Bessel function in
-\eqref{eq:O_time} as $\J_0\p{\tau}^2\approx1-\tau^2/2+O\p{\tau^4}$ and
-simplify the resulting expression to get
+At ``short'' times $t$ for which $tJ\sin\p{I\phi}\ll1$, we can expand
+the Bessel function in \eqref{eq:O_time} as
+$\J_0\p{\tau}^2\approx1-\tau^2/2+O\p{\tau^4}$ to get
+\begin{align}
+  \bk{\Psi|\O\p{t}|\Psi}
+  \approx 1 - \f12 \p{tJ}^2 \sum_{\mu,\nu}
+  \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2 \sin\p{\phi_{\mu\nu}^-}^2.
+\end{align}
+In the limit of a small angle $\phi$ for which $I\phi\ll1$, we can
+further simplify
 \begin{align}
@@
   \var\p{X} \equiv \bk{\psi|X^2|\psi} - \bk{\psi|X|\psi}^2.
+  \label{eq:small_time_angle}
 \end{align}
-Figure \ref{fig:limiting_vals} shows several values of
-$\var\p{s_\z/I}$ for different $n$.
+Note that the result in \eqref{eq:small_time_angle} is valid for all
+$tJ\ll\p{I\phi}^{-1}$, which may be a long time given the small-angle
+assumption $I\phi\ll1$.  Figure \ref{fig:limiting_vals} shows several
+values of $\var\p{s_\z}$ for different $n$.
+
+\begin{figure}
+  \centering
+  \includegraphics{figures/oscillations/const_vals.pdf}
+  \includegraphics{figures/oscillations/var_sz_vals.pdf}
+  \caption{Several values of $r$ and $\var\p{s_\z/I}$ for different
+    $n$.  Computed with respect to the state $\ket\psi$ in
+    \eqref{eq:double_state}.}
+  \label{fig:limiting_vals}
+\end{figure}
 

### 2020-07-30 sun_phases/scripts/plot_limiting_vals.py

@@ -74,7 +74,7 @@ plt.figure(figsize = figsize)
 plt.plot(n_vals, var_sz_vals/S_vals**2, "k.")
 plt.gca().set_ylim(bottom = 0)
 plt.xlabel("$n$")
-plt.ylabel(r"$\mathrm{var}(s_{\mathrm{z}}/I)$")
+plt.ylabel(r"$\mathrm{var}(s_{\mathrm{z}})/I^2$")
 plt.tight_layout()
 plt.savefig("../figures/oscillations/var_sz_vals.pdf")
 

### 2020-07-30 sun_phases/scripts/plot_limiting_vals.py
changed convention for one of the limiting case figures

@@ -16,7 +16,6 @@ plt.rcParams.update(params)
 ##############################
 
 n_vals = np.arange(2,21,2)
-S_vals = (n_vals-1)/2
 def spin_vals(n):
     S = (n-1)/2
     return np.linspace(-S,S,n)
@@ -58,7 +57,6 @@ plt.xlabel("$n$")
 plt.ylabel("$r$")
 plt.tight_layout()
 plt.savefig("../figures/oscillations/const_vals.pdf")
-plt.close("all")
 
 ##############################
 
@@ -71,10 +69,9 @@ def var_sz(n):
 var_sz_vals = [ var_sz(n) for n in n_vals ]
 
 plt.figure(figsize = figsize)
-plt.plot(n_vals, var_sz_vals/S_vals**2, "k.")
-plt.gca().set_ylim(bottom = 0)
+plt.plot(n_vals, var_sz_vals, "k.")
 plt.xlabel("$n$")
-plt.ylabel(r"$\mathrm{var}(s_{\mathrm{z}})/I^2$")
+plt.ylabel(r"$\mathrm{var}(s_{\mathrm{z}})$")
 plt.tight_layout()
 plt.savefig("../figures/oscillations/var_sz_vals.pdf")
 

### 2020-08-05 sun_phases/multilevel_spin_notes.tex
fixed up section on nuclear spin drive operators

--- 2020-07-22 collective_sun/multilevel_spin_notes.tex
@@
 \end{align}
-The effective driving amplitudes $\Omega_{LM}$ can be tuned through
-the bare driving amplitudes $\Omega_m$, and satisfy
-\begin{align}
-  \Omega_{0,0} &\propto \Omega_+^2 + \Omega_-^2 + \Omega_0^2,
-  &
-  \Omega_{1,0} &\propto \Omega_+^2 - \Omega_-^2,
-  &
-  \Omega_{2,0} &\propto \Omega_+^2 + \Omega_-^2 - 2\Omega_0^2,
-  \\
-  \Omega_{1,1} &\propto \Omega_0 \p{\Omega_+ + \Omega_-},
-  &
-  \Omega_{2,1} &\propto \Omega_0 \p{\Omega_+ - \Omega_-},
-  &
-  \Omega_{2,2} &\propto \Omega_+ \Omega_-.
-\end{align}
-Note that $\D_{0,0}$ is proportional to the identity operator, so the
-amplitude $\Omega_{0,0}$ can be neglected without consequence.  If
-$\Omega_0=\Omega_+=\Omega_-=\xi_1\Omega$, then all remaining
-amplitudes except $\Omega_{1,1}$ and $\Omega_{2,2}$ vanish, and the
-Raman drive becomes
-\begin{align}
-  H_{\t{Raman}}^{(+)} \equiv \f{\Omega^2}{\Delta}
-  \p{\sqrt{2}\, \col{S_\x} + \col{S_\x^2 - S_\y^2}},
-\end{align}
-where the single-spin operators $S_\alpha$ are related to the drive
-operators $D_{LM}$ through the relations in \eqref{eq:spin_ops},
-\eqref{eq:spin_ops_2_0}, and \eqref{eq:spin_ops_2_12}; and $\col{X}$
-is the collective version of $X$.  If instead
-$\Omega_0=\pm\Omega_\pm=\xi_1\Omega$, then all amplitudes except
-$\Omega_{2,1}$ and $\Omega_{2,2}$ vanish, and the drive becomes
-\begin{align}
-  H_{\t{Raman}}^{(-)} \equiv \f{\Omega^2}{\Delta}
-  \p{\sqrt{2}\, \sp{\col{S_\z S_\x+ S_\x S_\z}}
-    - \sp{\col{S_\x^2 - S_\y^2}}}.
-\end{align}
+Writing $H_{\t{Raman}} = \sum_j H_{\t{Raman}}^{\t{single},j}$, where
+$H_{\t{Raman}}^{\t{single},j}$ denotes the action of
+$H_{\t{Raman}}^{\t{single}}$ on spin $j$, we can expand
+\begin{align}
+  H_{\t{Raman}}^{\t{single}} \times \Delta
+  = h_+ h_- S_\z + h_0 h_+ S_\x
+  + h_0 h_- \p{S_\z S_\x  + S_\x S_\z}
+  + \p{h_0^2 - h_-^2} S_\x^2 + \p{h_0^2 - h_+^2} S_\y^2,
+\end{align}
+where
+$\p{S_\z,S_\x,S_\y} \equiv \xi_1\times\p{D_{1,0},D_{1,1},D_{1,-1}}$ is
+the standard SU(2) subalgebra of the operators on a single $n$-level
+spin, and
+\begin{align}
+  h_0 \equiv \xi_1^{-1}\, \Omega_0,
+  &&
+  h_\pm \equiv \xi_1^{-1}\, \f{\Omega_+\pm\Omega_-}{\sqrt{2}}.
+\end{align}
+Tuning the amplitudes $\Omega_0,\Omega_+,\Omega_-$, or equivalently
+$h_0,h_+,h_-$, enables us to implement various drive Hamiltonians,
+several of which are shown in Table \ref{tab:drives}.
+
+\begin{table}
+  \centering
+  \caption{Drive Hamiltonians that can be implemented with various
+    amplitude-matching conditions.}
+  \begin{tabular}{c|c|c}
+    Condition & $\p{h_0,h_+,h_-}/h$
+    & $H_{\t{Raman}}^{\t{single}}/\p{h^2/\Delta}$ \\ \hline\hline
+    $\Omega_0=\Omega_+=+\Omega_-$ & $\p{1,\sqrt{2},0}$
+    & $\sqrt{2} S_\x + S_\x^2 - S_\y^2$ \\ \hline
+    $\Omega_0=\Omega_+=-\Omega_-$ & $\p{1,0,\sqrt{2}}$
+    & $\sqrt{2}\p{S_\z S_\x + S_\x S_\z} - \p{S_\x^2-S_\y^2}$ \\ \hline
+    $\Omega_+=\Omega_-=0$ & $\p{1,0,0}$
+    & $S_\x^2 + S_\y^2$ \\ \hline
+    $\Omega_0=\Omega_-=0$ & $\p{0,1,+1}$
+    & $+S_\z - \p{S_\x^2 + S_\y^2}$ \\ \hline
+    $\Omega_0=\Omega_+=0$ & $\p{0,1,-1}$
+    & $-S_\z - \p{S_\x^2 + S_\y^2}$
+  \end{tabular}
+  \label{tab:drives}
+\end{table}
 

### 2020-08-06 sun_phases/sun_dynamics_notes.tex
changed convention for one of the limiting case figures
rearranged some notes and changed some minor conventions

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Non-interacting limit of a realistic 1D model}
+\section{Initial state}
 
-Consider the single-particle Hamiltonian
-\begin{align}
-  H = -\f{J}{2} \sum_{\mu,q} \cos\p{q + \mu\phi} s_{\mu\mu q},
-\end{align}
-where $q$ is a quasi-momentum (in units with lattice spacing $a=1$);
-$\mu\in\set{-I,-I+1,\cdots,I}$ is a nuclear spin projection onto a
-quantization axis, with $I\equiv\p{n-1}/2$ the total nuclear spin;
-$\phi$ is a spin-orbit-coupling angle, and $J$ is the single-particle
-bandwidth (proportional to the tunneling rate, $J/4$).  We wish to
-compute an expectation value of the operator
-\begin{align}
-  \O\p{t}
-  \equiv e^{\i t H} \p{\f{\v S^\dag \c\v S}{N^2}} e^{-\i t H}
-  = \f1{N^2} \sum_{\mu,\nu,p,q} s_{\mu\nu p} s_{\nu\mu q} \times
-  e^{\i tJ\p{\eta_{\mu\nu p}-\eta_{\mu\nu q}}},
-\end{align}
-where
-\begin{align}
-  \eta_{\mu\nu p}
-  \equiv -\f12 \sp{\cos\p{p+\mu\phi} - \cos\p{p+\nu\phi}}
-  = \sin\p{\phi_{\mu\nu}^-} \sin\p{p+\phi_{\mu\nu}^+},
-  &&
-  \phi_{\mu\nu}^\pm \equiv \f{\mu\pm\nu}{2}\,\phi.
-\end{align}
-If $\ket\Psi=\ket\psi^{\otimes N}$, then
-\begin{align}
-  \bk{\Psi|\O\p{t}|\Psi}
-  &= \f1{N^2} \sum_{\mu,\nu} \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2
-  \sum_{p\ne q} e^{\i tJ\p{\eta_{\mu\nu p}-\eta_{\mu\nu q}}} + \f{n}{N} \\
-  &= \sum_{\mu,\nu} \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2
-  \abs{\xi_{\mu\nu}\p{\phi}}^2 + \f{n-1}{N},
-\end{align}
-where
-\begin{align}
-  \xi_{\mu\nu}\p{\phi} \equiv \f1N \sum_q e^{\i tJ\eta_{\mu\nu q}}
-  = \f1N \sum_q e^{\i tJ\sin\p{\phi_{\mu\nu}^-} \sin\p{q+\phi_{\mu\nu}^+}}.
-\end{align}
-As $N\to\infty$, this sum is well approximated by an integral:
-\begin{align}
-  \xi_{\mu\nu}\p{\phi} \approx \f1{2\pi} \int_{-\pi}^\pi \dd\theta\,
-  e^{\i tJ\sin\p{\phi_{\mu\nu}^-} \sin\p{\theta+\phi_{\mu\nu}^+}}
-  = \J_0\sp{tJ\sin\p{\phi_{\mu\nu}^-}},
-\end{align}
-where $\J_0$ is the zero-order Bessel function of the first kind.  To
-leading order in $N\gg1$, we thus find that
-\begin{align}
-  \bk{\Psi|\O\p{t}|\Psi}
-  \approx \sum_{\mu,\nu} \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2
-  \J_0\sp{tJ\sin\p{\phi_{\mu\nu}^-}}^2.
-  \label{eq:O_time}
-\end{align}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Initial state}
-
-Up to an inconsequential choice of gauge, the initial state that we
-can prepare has the form
+We can prepare an initial product state
+$\ket\Psi\equiv\ket\psi^{\otimes N}$ with
 \begin{align}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Non-interacting limit of a realistic 1D model}
+
+We consider the single-particle Hamiltonian
+\begin{align}
+  H = -\f{J}{2} \sum_{\mu,q} \cos\p{q + \mu\phi} s_{\mu\mu q},
+\end{align}
+where $q$ is a quasi-momentum (in units with lattice spacing $a=1$);
+$\mu\in\set{-I,-I+1,\cdots,I}$ is a nuclear spin projection onto a
+quantization axis, with $I\equiv\p{n-1}/2$ the total nuclear spin;
+$\phi$ is a spin-orbit-coupling angle, and $J$ is the single-particle
+bandwidth (proportional to the tunneling rate, $J/4$).  As a dynamical
+order parameter, we consider the operator
+\begin{align}
+  \O\p{t}
+  \equiv e^{\i t H} \p{\f{\v S^\dag \c\v S}{N^2}} e^{-\i t H}
+  = \f1{N^2} \sum_{\mu,\nu,p,q} s_{\mu\nu p} s_{\nu\mu q} \times
+  e^{\i tJ\p{\eta_{\mu\nu p}-\eta_{\mu\nu q}}},
+\end{align}
+where
+\begin{align}
+  \eta_{\mu\nu p}
+  \equiv -\f12 \sp{\cos\p{p+\mu\phi} - \cos\p{p+\nu\phi}}
+  = \sin\p{\phi_{\mu\nu}^-} \sin\p{p+\phi_{\mu\nu}^+},
+  &&
+  \phi_{\mu\nu}^\pm \equiv \f{\mu\pm\nu}{2}\,\phi.
+\end{align}
+Specifically, we are interested in the expectation value
+\begin{align}
+  \bk{\Psi|\O\p{t}|\Psi}
+  &= \f1{N^2} \sum_{\mu,\nu} \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2
+  \sum_{p\ne q} e^{\i tJ\p{\eta_{\mu\nu p}-\eta_{\mu\nu q}}} + \f{n}{N} \\
+  &= \sum_{\mu,\nu} \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2
+  \abs{\xi_{\mu\nu}\p{\phi}}^2 + \f{n-1}{N},
+\end{align}
+where
+\begin{align}
+  \xi_{\mu\nu}\p{\phi} \equiv \f1N \sum_q e^{\i tJ\eta_{\mu\nu q}}
+  = \f1N \sum_q e^{\i tJ\sin\p{\phi_{\mu\nu}^-} \sin\p{q+\phi_{\mu\nu}^+}}.
+\end{align}
+As $N\to\infty$, this sum is well approximated by an integral:
+\begin{align}
+  \xi_{\mu\nu}\p{\phi} \approx \f1{2\pi} \int_{-\pi}^\pi \dd\theta\,
+  e^{\i tJ\sin\p{\phi_{\mu\nu}^-} \sin\p{\theta+\phi_{\mu\nu}^+}}
+  = \J_0\sp{tJ\sin\p{\phi_{\mu\nu}^-}},
+\end{align}
+where $\J_0$ is the zero-order Bessel function of the first kind.  To
+leading order in $N\gg1$, we thus find that
+\begin{align}
+  \bk{\Psi|\O\p{t}|\Psi}
+  \approx \sum_{\mu,\nu} \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2
+  \J_0\sp{tJ\sin\p{\phi_{\mu\nu}^-}}^2.
+  \label{eq:O_time}
+\end{align}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-We now return to the order parameter $\O\p{t}$ in \eqref{eq:O_time}.
-The constant (time independent) terms in this expression have
+The constant (time independent) terms in \eqref{eq:O_time} have
 $\mu=\nu$ (in which case all $\phi_{\mu\nu}^-=0$), so up to $O(1/N)$
@@
   \bk{\Psi|\O\p{t}|\Psi}
-  \approx 1 - \f14 \var\p{s_\z} \p{tJ\phi}^2,
-  &&
-  \var\p{X} \equiv \bk{\psi|X^2|\psi} - \bk{\psi|X|\psi}^2.
+  \approx 1 - \f14 \var\p{Z} \p{JI\phi}^2 \times t^2,
   \label{eq:small_time_angle}
 \end{align}
+where
+\begin{align}
+  \var\p{Z}
+  \equiv \bk{\psi|Z^2|\psi} - \bk{\psi|Z|\psi}^2,
+  &&
+  Z \equiv \sum_{\mu=-I}^I \f{\mu}{I} \op{\mu}.
+\end{align}
 Note that the result in \eqref{eq:small_time_angle} is valid for all
-$tJ\ll\p{I\phi}^{-1}$, which may be a long time given the small-angle
+$t\ll\p{JI\phi}^{-1}$, which may be a long time given the small-angle
 assumption $I\phi\ll1$.  Figure \ref{fig:limiting_vals} shows several
-values of $\var\p{s_\z}$ for different $n$.
+values of $\var\p{Z}$ for different $n$.
 
@@
   \includegraphics{figures/oscillations/const_vals.pdf}
-  \includegraphics{figures/oscillations/var_sz_vals.pdf}
-  \caption{Several values of $r$ and $\var\p{s_\z/I}$ for different
-    $n$.  Computed with respect to the state $\ket\psi$ in
+  \includegraphics{figures/oscillations/var_z_vals.pdf}
+  \caption{Several values of $r$ and $\var\p{Z}$ for different $n$.
+    Computed with respect to the state $\ket\psi$ in
     \eqref{eq:double_state}.}
@@
 \end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Strongly-interacting limit of a realistic 1D model}
+
+We now consider the opposite, strongly-interacting limit of a 1D
+model,
 

### 2020-08-06 sun_phases/scripts/plot_limiting_vals.py

@@ -60,18 +60,26 @@ plt.savefig("../figures/oscillations/const_vals.pdf")
 
 ##############################
 
-def var_sz(n):
+def var_Z(n):
     _state = state(n)
-    _sz = spin_vals(n)
-    val_sz_sz = _state.conj() @ ( _sz**2 * _state )
-    val_sz = _state.conj() @ ( _sz * _state )
-    return ( val_sz_sz - val_sz**2 ).real
-var_sz_vals = [ var_sz(n) for n in n_vals ]
+    S = (n-1)/2
+    Z = spin_vals(n) / S
+    val_Z_Z = _state.conj() @ ( Z**2 * _state )
+    val_Z = _state.conj() @ ( Z * _state )
+    return ( val_Z_Z - val_Z**2 ).real
+var_Z_vals = np.array([ var_Z(n) for n in n_vals ])
 
 plt.figure(figsize = figsize)
-plt.plot(n_vals, var_sz_vals, "k.")
+plt.plot(n_vals, var_Z_vals, "k.")
 plt.xlabel("$n$")
-plt.ylabel(r"$\mathrm{var}(s_{\mathrm{z}})$")
+plt.ylabel(r"$\mathrm{var}(Z)$")
 plt.tight_layout()
-plt.savefig("../figures/oscillations/var_sz_vals.pdf")
+plt.savefig("../figures/oscillations/var_z_vals.pdf")
 
+
+plt.figure(figsize = figsize)
+plt.plot(n_vals, np.sqrt(var_Z_vals), "k.")
+plt.xlabel("$n$")
+plt.ylabel(r"$\mathrm{std}(Z)$")
+plt.tight_layout()
+plt.savefig("../figures/oscillations/std_z_vals.pdf")

### 2020-08-06 common_codes/dicke_methods.py
add more drive Hamiltonians

@@ -100,7 +100,7 @@ try:
         return cmocean.cm.amp(color_vals)
 except ModuleNotFoundError:
     def sphere_cmap(color_vals):
-        return plt.cm.get_cmap("inferno")(1-color_vals)
+        return plt.cm.get_cmap("inferno")(color_vals)
 
 def plot_dicke_state(state, grid_size = 101, single_sphere = True, figsize = None,
                      rasterized = True):

### 2020-08-06 sun_phases/sun_dynamics_notes.tex
add more drive Hamiltonians
allow simulating with different initial states

@@
 \newcommand{\z}{\text{z}}
+\newcommand{\X}{\text{X}}
+\newcommand{\DS}{\text{DS}}
 
@@
 \newcommand{\V}{\mathcal{V}}
-\newcommand{\X}{\mathcal{X}}
 \newcommand{\Z}{\mathcal{Z}}
@@
 \end{align}
-where $\mu,\nu\in\ZZ_n$ index orthogonal states of a single spin;
-$i,j\in\ZZ_N$ index one of $N$ spins; $h^{\mu\nu i}_{\rho\sigma j}$
-and $\epsilon_{\mu\nu i}$ are scalars; and
-$s_{\mu\nu i}\coloneqq\op{\mu}{\nu}_i$ is a transition operator for
-spin $i$.  If interactions are SU($n$) symmetric then
+where $\mu,\nu\in\set{-I,-I+1,\cdots,I}$ with $I\equiv\p{n-1}/2$ index
+orthogonal states of an $n$-level spin; $i,j\in\ZZ_N$ index one of $N$
+spins; $h^{\mu\nu i}_{\rho\sigma j}$ and $\epsilon_{\mu\nu i}$ are
+scalars; and $s_{\mu\nu i}\coloneqq\op{\mu}{\nu}_i$ is a transition
+operator for spin $i$.  If interactions are SU($n$) symmetric then
 $h^{\mu\nu i}_{\rho\sigma k} = h_{ik}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Initial state}
+\section{Initial states}
 
-We can prepare an initial product state
-$\ket\Psi\equiv\ket\psi^{\otimes N}$ with
+We can essentially prepare two types of initial product states
+$\ket\Psi\equiv\ket\psi^{\otimes N}$.  First and foremost, we can
+prepare arbitrary polarized states defined by the polar and azimuthal
+angles $\theta,\phi$:
 \begin{align}
-  \ket\psi = \f1{\sqrt{2}}
-  \p{e^{\i\beta} \ket{\theta,\alpha}
-    + e^{-\i\beta} \ket{\theta,-\alpha}},
-  &&
-  \ket{\theta,\alpha} \equiv e^{i\alpha s_\z} e^{i\theta s_\y} \ket\up,
-  \label{eq:double_state}
+  \ket{\theta,\phi}
+  \equiv e^{i\phi s_\z} e^{i\theta s_\y} \ket\up
+  = \sum_\mu \zeta_\mu\p{\theta} e^{-\i\mu\phi} \ket\mu,
+  \label{eq:polarized_state}
 \end{align}
-where $\theta,\alpha$ are respectively polar and azimuthal angles of
-the polarized state $\ket{\theta,\alpha}$ of a single spin; the spin
-operators $s_\z,s_\y$ generate rotations about the $z,y$ axes, with
-e.g.~$s_\z\equiv\sum_\mu \mu\op\mu$; and $\ket\up$ is a state
-polarized along the $z$ axis.  More specifically, the state that we
-can prepare has
-\begin{align}
-  \theta = \f{\pi}{2} + \arcsin\p{\f13},
-  &&
-  \alpha = \f{\pi}{3},
-  &&
-  \beta = -\f{2\pi}{3}\, I.
-\end{align}
-In order to compute expectation values with respect to $\ket\psi$, we
-expand\footnote{The expansion in \eqref{eq:polarized_proj} is acquired
+where $\ket\up\equiv\ket{I}$ is a state polarized along the $z$ axis,
+and\footnote{The expansion of $\ket{\theta,\phi}$ in
+  \eqref{eq:polarized_state} and \eqref{eq:polarized_proj} is acquired
   by representing an $n$-level spin by $n-1$ qubits restricted to the
@@
 \begin{align}
-  \ket{\theta,\alpha} = \sum_\mu \chi_\mu e^{-\i\mu\alpha} \ket\mu,
-  &&
-  \chi_\mu \equiv { 2I \choose I+\mu }^{1/2}
-  \cos\p{\f{\theta}{2}}^{I+\mu} \sin\p{\f{\theta}{2}}^{I-\mu},
+  \zeta_\mu\p{\theta} \equiv { 2I \choose I+\mu }^{1/2}
+  \cos\p{\f{\theta}{2}}^{I+\mu} \sin\p{\f{\theta}{2}}^{I-\mu}.
   \label{eq:polarized_proj}
 \end{align}
-which implies that
+For even $n$, we can also prepare the following superposition of
+polarized states:
 \begin{align}
-  \ket\psi = \sum_\mu \chi_\mu \times \sqrt{2}\,
-  \cos\p{\mu\alpha-\beta} \ket\mu.
+  \ket{\DS}
+  \equiv \f1{\sqrt{2}}
+  \p{e^{\i\gamma} \ket{\alpha,\beta}
+    + e^{-\i\gamma} \ket{\alpha,-\beta}}
+  = \sum_\mu \zeta_\mu\p{\alpha}
+  \times \sqrt{2}\, \cos\p{\mu\beta-\gamma} \ket\mu
+  \label{eq:double_state}
+\end{align}
+with
+\begin{align}
+  \alpha = \f{\pi}{2} + \arcsin\p{\f13},
+  &&
+  \beta = \f{\pi}{3},
+  &&
+  \gamma = -\f{2\pi}{3}\, I.
 \end{align}
@@
 \begin{align}
-  H = -\f{J}{2} \sum_{\mu,q} \cos\p{q + \mu\phi} s_{\mu\mu q},
+  H = -\f{J}{2} \sum_{\mu,q} \cos\p{q + \mu\xi} s_{\mu\mu q},
 \end{align}
@@
 quantization axis, with $I\equiv\p{n-1}/2$ the total nuclear spin;
-$\phi$ is a spin-orbit-coupling angle, and $J$ is the single-particle
+$\xi$ is a spin-orbit-coupling angle, and $J$ is the single-particle
 bandwidth (proportional to the tunneling rate, $J/4$).  As a dynamical
@@
   \eta_{\mu\nu p}
-  \equiv -\f12 \sp{\cos\p{p+\mu\phi} - \cos\p{p+\nu\phi}}
-  = \sin\p{\phi_{\mu\nu}^-} \sin\p{p+\phi_{\mu\nu}^+},
+  \equiv -\f12 \sp{\cos\p{p+\mu\xi} - \cos\p{p+\nu\xi}}
+  = \sin\p{\xi_{\mu\nu}^-} \sin\p{p+\xi_{\mu\nu}^+},
   &&
-  \phi_{\mu\nu}^\pm \equiv \f{\mu\pm\nu}{2}\,\phi.
+  \xi_{\mu\nu}^\pm \equiv \f{\mu\pm\nu}{2}\,\xi.
 \end{align}
@@
   &= \sum_{\mu,\nu} \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2
-  \abs{\xi_{\mu\nu}\p{\phi}}^2 + \f{n-1}{N},
+  \abs{\chi_{\mu\nu}\p{\xi}}^2 + \f{n-1}{N},
 \end{align}
@@
 \begin{align}
-  \xi_{\mu\nu}\p{\phi} \equiv \f1N \sum_q e^{\i tJ\eta_{\mu\nu q}}
-  = \f1N \sum_q e^{\i tJ\sin\p{\phi_{\mu\nu}^-} \sin\p{q+\phi_{\mu\nu}^+}}.
+  \chi_{\mu\nu}\p{\xi} \equiv \f1N \sum_q e^{\i tJ\eta_{\mu\nu q}}
+  = \f1N \sum_q e^{\i tJ\sin\p{\xi_{\mu\nu}^-} \sin\p{q+\xi_{\mu\nu}^+}}.
 \end{align}
@@
 \begin{align}
-  \xi_{\mu\nu}\p{\phi} \approx \f1{2\pi} \int_{-\pi}^\pi \dd\theta\,
-  e^{\i tJ\sin\p{\phi_{\mu\nu}^-} \sin\p{\theta+\phi_{\mu\nu}^+}}
-  = \J_0\sp{tJ\sin\p{\phi_{\mu\nu}^-}},
+  \chi_{\mu\nu}\p{\xi} \approx \f1{2\pi} \int_{-\pi}^\pi \dd\theta\,
+  e^{\i tJ\sin\p{\xi_{\mu\nu}^-} \sin\p{\theta+\xi_{\mu\nu}^+}}
+  = \J_0\sp{tJ\sin\p{\xi_{\mu\nu}^-}},
 \end{align}
@@
   \approx \sum_{\mu,\nu} \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2
-  \J_0\sp{tJ\sin\p{\phi_{\mu\nu}^-}}^2.
+  \J_0\sp{tJ\sin\p{\xi_{\mu\nu}^-}}^2.
   \label{eq:O_time}
@@
 The constant (time independent) terms in \eqref{eq:O_time} have
-$\mu=\nu$ (in which case all $\phi_{\mu\nu}^-=0$), so up to $O(1/N)$
+$\mu=\nu$ (in which case all $\xi_{\mu\nu}^-=0$), so up to $O(1/N)$
 corrections
@@
   \bk{\Psi|\O_{\t{const}}|\Psi}
-  \approx r \equiv \sum_\mu \abs{\bk{\mu|\psi}}^4,
+  \approx r_\psi \equiv \sum_\mu \abs{\bk{\mu|\psi}}^4,
   \label{eq:O_const}
@@
 which is sometimes called the inverse participation ratio of
-$\ket\psi$ with respect to the basis $\set{\ket\mu}$.  Figure
-\ref{fig:limiting_vals} shows several values of $r$ for different $n$.
+$\ket\psi$ with respect to the basis $\set{\ket\mu}$ of states with
+definite projection onto the $z$ axis.  Figure \ref{fig:limiting_vals}
+shows several values of $r_\psi$ for different $n$.
 
-At ``short'' times $t$ for which $tJ\sin\p{I\phi}\ll1$, we can expand
+At ``short'' times $t$ for which $tJ\sin\p{I\xi}\ll1$, we can expand
 the Bessel function in \eqref{eq:O_time} as
@@
   \approx 1 - \f12 \p{tJ}^2 \sum_{\mu,\nu}
-  \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2 \sin\p{\phi_{\mu\nu}^-}^2.
+  \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2 \sin\p{\xi_{\mu\nu}^-}^2.
 \end{align}
-In the limit of a small angle $\phi$ for which $I\phi\ll1$, we can
+In the limit of a small angle $\xi$ for which $I\xi\ll1$, we can
 further simplify
@@
   \bk{\Psi|\O\p{t}|\Psi}
-  \approx 1 - \f14 \var\p{Z} \p{JI\phi}^2 \times t^2,
+  \approx 1 - \f14 \var_\psi\p{Z} \p{JI\xi}^2 \times t^2,
   \label{eq:small_time_angle}
@@
 \begin{align}
-  \var\p{Z}
-  \equiv \bk{\psi|Z^2|\psi} - \bk{\psi|Z|\psi}^2,
+  \var_\psi\p{Z} \equiv \bk{\psi|Z^2|\psi} - \bk{\psi|Z|\psi}^2,
   &&
-  Z \equiv \sum_{\mu=-I}^I \f{\mu}{I} \op{\mu}.
+  Z \equiv \sum_\mu \f{\mu}{I} \op{\mu}.
 \end{align}
 Note that the result in \eqref{eq:small_time_angle} is valid for all
-$t\ll\p{JI\phi}^{-1}$, which may be a long time given the small-angle
-assumption $I\phi\ll1$.  Figure \ref{fig:limiting_vals} shows several
-values of $\var\p{Z}$ for different $n$.
+$t\ll\p{JI\xi}^{-1}$, which may be a long time given the small-angle
+assumption $I\xi\ll1$.  Figure \ref{fig:limiting_vals} shows several
+values of $\var_\psi\p{Z}$ for different $n$.
 
@@
   \centering
-  \includegraphics{figures/oscillations/const_vals.pdf}
-  \includegraphics{figures/oscillations/var_z_vals.pdf}
-  \caption{Several values of $r$ and $\var\p{Z}$ for different $n$.
-    Computed with respect to the state $\ket\psi$ in
-    \eqref{eq:double_state}.}
+  \includegraphics{figures/oscillations/limiting_vals.pdf}
+  \caption{Several values of $r_\psi$ and $\var_\psi\p{Z}$ for
+    different $n$.  Computed with respect to the states
+    $\ket\X\equiv\ket{\pi/2,0}$ and $\ket{\DS}$, defined in
+    \eqref{eq:polarized_state} and \eqref{eq:double_state}.}
   \label{fig:limiting_vals}

### 2020-08-06 sun_phases/scripts/plot_bosons.py

@@ -16,9 +16,9 @@ color_map = "inferno"
 # plotting parameters
 log10_tun_vals = np.linspace(-2,1,13)
 soc_frac_vals = np.linspace(0,1,11)
+plot_peaks = 100
 freq_num = 1000
 freq_scale = 2
-plot_peaks = 100
 
 data_dir = "../data/oscillations/"
 fig_dir = "../figures/oscillations/"

### 2020-08-06 sun_phases/scripts/plot_drive_states.py

@@ -81,20 +81,20 @@ if any( np.isclose(tau % 1, [0,1/2,1]) ):
 
 if any( np.isclose(tau % 1, [1/4,3/4]) ):
     if dim % 2 == 0:
-        theta_1 = np.pi/2 + np.arcsin(1/3)
-        theta_2 = np.pi/2 + np.arcsin(1/3)
-        alpha = np.pi/3
-        beta = -(dim-1)/3 * np.pi
+        alpha_1 = np.pi/2 + np.arcsin(1/3)
+        alpha_2 = np.pi/2 + np.arcsin(1/3)
+        beta = np.pi/3
+        gamma = -(dim-1)/3 * np.pi
 
     else:
-        theta_1 = np.pi
-        theta_2 = np.pi/2 - np.arcsin(1/3)
-        alpha = 0
-        beta = (dim-2)/4 * np.pi
-        if np.isclose(tau,3/4): beta *= -1
+        alpha_1 = np.pi
+        alpha_2 = np.pi/2 - np.arcsin(1/3)
+        beta = 0
+        gamma = (dim-2)/4 * np.pi
+        if np.isclose(tau,3/4): gamma *= -1
 
-    guess = ( polarized_state(theta_1, +alpha) * np.exp(+1j*beta) +
-              polarized_state(theta_2, -alpha) * np.exp(-1j*beta) )
+    guess = ( polarized_state(alpha_1, +beta) * np.exp(+1j*gamma) +
+              polarized_state(alpha_2, -beta) * np.exp(-1j*gamma) )
 
 badness = infidelity(guess)
 

### 2020-08-06 sun_phases/scripts/plot_limiting_vals.py

@@ -6,9 +6,9 @@ import scipy.special
 
 from dicke_methods import coherent_spin_state_angles
 
-figsize = (3,2)
+figsize = (4,1.5)
 
-params = { "font.size" : 12,
+params = { "font.size" : 10,
            "text.usetex" : True,
            "text.latex.preamble" : [ r"\usepackage{physics}" ]}
 plt.rcParams.update(params)
@@ -16,70 +16,58 @@ plt.rcParams.update(params)
 ##############################
 
 n_vals = np.arange(2,21,2)
+n_ticks = np.arange(4,21,4)
+
+I_vals = (n_vals-1)/2
 def spin_vals(n):
-    S = (n-1)/2
-    return np.linspace(-S,S,n)
+    I = (n-1)/2
+    return np.linspace(-I,I,n)
 
-theta = np.pi/2 + np.arcsin(1/3)
-alpha = np.pi/3
-def beta(n):
-    return -(n-1)/3 * np.pi
-def state(n):
-    return ( coherent_spin_state_angles(theta, +alpha, n-1) * np.exp(+1j*beta(n)) +
-             coherent_spin_state_angles(theta, -alpha, n-1) * np.exp(-1j*beta(n)) ) / np.sqrt(2)
+alpha = np.pi/2 + np.arcsin(1/3)
+beta = np.pi/3
+def gamma(n): return -(n-1)/3 * np.pi
 
-##############################
-
-def chi(m,n):
-    S = (n-1)/2
-    return ( np.sqrt(scipy.special.binom(2*S,S+m)) *
-             np.cos(theta/2)**(S+m) * np.sin(theta/2)**(S-m) )
-def proj(m,n):
-    return chi(m,n) * np.sqrt(2) * np.cos(m*alpha-beta(n))
-
-for n in n_vals:
-    assert(np.allclose( state(n), [ proj(m,n) for m in spin_vals(n) ] ))
+states = [ "X", "DS" ]
+def state_vec(n, state):
+    if state == "X":
+        return coherent_spin_state_angles(np.pi/2, 0, n-1)
+    elif state == "DS":
+        state_p = coherent_spin_state_angles(alpha, +beta, n-1) * np.exp(+1j*gamma(n))
+        state_m = coherent_spin_state_angles(alpha, -beta, n-1) * np.exp(-1j*gamma(n))
[…]

### 2020-08-12 sun_phases/sun_dynamics_notes.tex
work out mean single-particle observables in the non-interacting limit

@@
 \end{align}
-where $\mu,\nu\in\set{-I,-I+1,\cdots,I}$ with $I\equiv\p{n-1}/2$ index
-orthogonal states of an $n$-level spin; $i,j\in\ZZ_N$ index one of $N$
-spins; $h^{\mu\nu i}_{\rho\sigma j}$ and $\epsilon_{\mu\nu i}$ are
-scalars; and $s_{\mu\nu i}\coloneqq\op{\mu}{\nu}_i$ is a transition
-operator for spin $i$.  If interactions are SU($n$) symmetric then
-$h^{\mu\nu i}_{\rho\sigma k} = h_{ik}
-\delta_{\mu\sigma}\delta_{\nu\rho}$, in which case
+where $\mu,\nu\in\set{I,I-1,\cdots,-I}$ with $I\equiv\p{n-1}/2$ index orthogonal states of an $n$-level spin; $i,j\in\ZZ_N$ index one of $N$ spins; $h^{\mu\nu i}_{\rho\sigma j}$ and $\epsilon_{\mu\nu i}$ are scalars; and $s_{\mu\nu i}\coloneqq\op{\mu}{\nu}_i$ is a transition operator for spin $i$.
+If interactions are SU($n$) symmetric then $h^{\mu\nu i}_{\rho\sigma k} = h_{ik} \delta_{\mu\sigma}\delta_{\nu\rho}$, in which case
 \begin{align}
@@
 
-We can write the Hamiltonian in \eqref{eq:spin} using Schwinger bosons
-as
+We can write the Hamiltonian in \eqref{eq:spin} using Schwinger bosons as
 \begin{align}
@@
 \end{align}
-where $b_{\mu i}$ annihilates a boson of type $\mu$ on site $i$.  The
-Heisenberg equations of motion for these operators are\footnote{The
-  Hamiltonian in \eqref{eq:spin} only defines
-  $h^{\mu\nu j}_{\rho\sigma k}$ for $j<k$.  To simplify expressions in
-  the remainder of this work, we therefore additionally define
-  $h^{\mu\nu k}_{\rho\sigma j}=h^{\rho\sigma j}_{\mu\nu k}$ and
-  $h^{\mu\nu i}_{\rho\sigma i}=0$.}
+where $b_{\mu i}$ annihilates a boson of type $\mu$ on site $i$.
+The Heisenberg equations of motion for these operators are\footnote{The Hamiltonian in \eqref{eq:spin} only defines $h^{\mu\nu j}_{\rho\sigma k}$ for $j<k$.
+  To simplify expressions in the remainder of this work, we therefore additionally define $h^{\mu\nu k}_{\rho\sigma j}=h^{\rho\sigma j}_{\mu\nu k}$ and $h^{\mu\nu i}_{\rho\sigma i}=0$.}
 \begin{align}
@@
 
-We can essentially prepare two types of initial product states
-$\ket\Psi\equiv\ket\psi^{\otimes N}$.  First and foremost, we can
-prepare arbitrary polarized states defined by the polar and azimuthal
-angles $\theta,\phi$:
+We can essentially prepare $N$-fold tensor products $\ket\Psi\equiv\ket\psi^{\otimes N}$ of three types of initial product states $\ket\psi$.
+First and foremost, we can prepare states $\ket{\mu}$ of definite spin projection $\mu=\set{I,I-1,\cdots,-I}$ onto a quantization axis.
+In addition, we can prepare arbitrary polarized states defined by the polar and azimuthal angles $\theta,\phi$:
 \begin{align}
@@
 \end{align}
-where $\ket\up\equiv\ket{I}$ is a state polarized along the $z$ axis,
-and\footnote{The expansion of $\ket{\theta,\phi}$ in
-  \eqref{eq:polarized_state} and \eqref{eq:polarized_proj} is acquired
-  by representing an $n$-level spin by $n-1$ qubits restricted to the
-  (Dicke) manifold of permutationally symmetric states.}
+where $\ket\up\equiv\ket{I}$ is a state polarized along the $z$ axis, and\footnote{The expansion of $\ket{\theta,\phi}$ in \eqref{eq:polarized_state} and \eqref{eq:polarized_proj} is acquired by representing an $n$-level spin by $n-1$ qubits restricted to the (Dicke) manifold of permutationally symmetric states.}
 \begin{align}
@@
 \end{align}
-For even $n$, we can also prepare the following superposition of
-polarized states:
+Finally, for even $n$ we can also prepare the following superposition of polarized states:
 \begin{align}
@@
 \end{align}
-where $q$ is a quasi-momentum (in units with lattice spacing $a=1$);
-$\mu\in\set{-I,-I+1,\cdots,I}$ is a nuclear spin projection onto a
-quantization axis, with $I\equiv\p{n-1}/2$ the total nuclear spin;
-$\xi$ is a spin-orbit-coupling angle, and $J$ is the single-particle
-bandwidth (proportional to the tunneling rate, $J/4$).  As a dynamical
-order parameter, we consider the operator
+where $q$ is a quasi-momentum (in units with lattice spacing $a=1$); $\mu\in\set{I,I-1,\cdots,-I}$ is a nuclear spin projection onto a quantization axis, with $I\equiv\p{n-1}/2$ the total nuclear spin; $\xi$ is a spin-orbit-coupling angle, and $J$ is the single-particle bandwidth (proportional to the tunneling rate, $J/4$).
+As dynamical order parameters, we consider the operators
 \begin{align}
-  \O\p{t}
-  \equiv e^{\i t H} \p{\f{\v S^\dag \c\v S}{N^2}} e^{-\i t H}
-  = \f1{N^2} \sum_{\mu,\nu,p,q} s_{\mu\nu p} s_{\nu\mu q} \times
-  e^{\i tJ\p{\eta_{\mu\nu p}-\eta_{\mu\nu q}}},
+  s_{\mu\nu} \equiv \f1N \sum_q s_{\mu\nu q},
+  &&
+  s^2 \equiv \v s^\dag\cdot \v s
+  = \f1{N^2} \sum_{\mu,\nu,p,q} s_{\mu\nu p} s_{\nu\mu q},
+\end{align}
+where $\v s_q\equiv\sum_{\mu,\nu} s_{\mu\nu q}\uv e_{\mu\nu}$ is a vector of normalized generators of $U(n)$ for spin $q$, and $\uv e_{\mu\nu}$ is a unit vector of $\RR^{n^2}$.
+In the Heisenberg picture, these operators at time $t$ are
+\begin{align}
+  s_{\mu\nu}\p{t} = \f1N \sum_q s_{\mu\nu q}
+  \times e^{\i t J \eta_{\mu\nu q}},
+  &&
+  s^2\p{t} = \f1{N^2} \sum_{\mu,\nu,p,q} s_{\mu\nu p} s_{\nu\mu q}
+  \times e^{\i tJ\p{\eta_{\mu\nu p}-\eta_{\mu\nu q}}},
 \end{align}
@@
 \begin{align}
-  \eta_{\mu\nu p}
-  \equiv -\f12 \sp{\cos\p{p+\mu\xi} - \cos\p{p+\nu\xi}}
+  \eta_{\mu\nu q}
+  \equiv -\f12 \sp{\cos\p{q+\mu\xi} - \cos\p{q+\nu\xi}}
   = \sin\p{\xi_{\mu\nu}^-} \sin\p{p+\xi_{\mu\nu}^+},
@@
 \end{align}
-Specifically, we are interested in the expectation value
+In particular, we are interested in the expectation values
 \begin{align}
-  \bk{\Psi|\O\p{t}|\Psi}
+  \bk{\Psi|s_{\mu\nu}\p{t}|\Psi}
+  = \f1N \bk{\psi\op{\mu}{\nu}\psi} \sum_q e^{\i t J \eta_{\mu\nu q}}
+  = \bk{\psi\op{\mu}{\nu}\psi} \chi_{\mu\nu}\p{\xi},
+\end{align}
+and
+\begin{align}
+  \bk{\Psi|s^2\p{t}|\Psi}
   &= \f1{N^2} \sum_{\mu,\nu} \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2
@@
 \end{align}
-where $\J_0$ is the zero-order Bessel function of the first kind.  To
-leading order in $N\gg1$, we thus find that
+where $\J_0$ is the zero-order Bessel function of the first kind.
+To leading order in $N\gg1$, we thus find that
 \begin{align}
-  \bk{\Psi|\O\p{t}|\Psi}
-  \approx \sum_{\mu,\nu} \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2
+  \bk{\Psi|s_{\mu\nu}\p{t}|\Psi}
+  &\approx \bk{\psi\op{\mu}{\nu}\psi}
+  \J_0\sp{tJ\sin\p{\xi_{\mu\nu}^-}},
+  \\
+  \bk{\Psi|s^2\p{t}|\Psi}
+  &\approx \sum_{\mu,\nu} \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2
   \J_0\sp{tJ\sin\p{\xi_{\mu\nu}^-}}^2.
-  \label{eq:O_time}
+  \label{eq:ops_time}
 \end{align}
@@
 
-The constant (time independent) terms in \eqref{eq:O_time} have
-$\mu=\nu$ (in which case all $\xi_{\mu\nu}^-=0$), so up to $O(1/N)$
-corrections
+The constant (time independent) terms in \eqref{eq:ops_time} have $\mu=\nu$ (in which case all $\xi_{\mu\nu}^-=0$), so up to $O(1/N)$ corrections
 \begin{align}
-  \bk{\Psi|\O_{\t{const}}|\Psi}
+  \bk{\Psi|s_{\mu\nu}^{\t{const}}|\Psi}
+  \approx \abs{\bk{\mu|\psi}}^2,
+  &&
+  \bk{\Psi|s^2_{\t{const}}|\Psi}
   \approx r_\psi \equiv \sum_\mu \abs{\bk{\mu|\psi}}^4,
-  \label{eq:O_const}
+  \label{eq:ops_const}
 \end{align}
-which is sometimes called the inverse participation ratio of
-$\ket\psi$ with respect to the basis $\set{\ket\mu}$ of states with
-definite projection onto the $z$ axis.  Figure \ref{fig:limiting_vals}
-shows several values of $r_\psi$ for different $n$.
+where the quantity $r_\psi$ is sometimes called the inverse participation ratio of $\ket\psi$ with respect to the basis $\set{\ket\mu}$ of states with definite projection onto the $z$ axis.
+Figure \ref{fig:limiting_vals} shows several values of $r_\psi$ for different $n$.
 
-At ``short'' times $t$ for which $tJ\sin\p{I\xi}\ll1$, we can expand
-the Bessel function in \eqref{eq:O_time} as
-$\J_0\p{\tau}^2\approx1-\tau^2/2+O\p{\tau^4}$ to get
+At ``short'' times $t$ for which $tJ\sin\p{I\xi}\ll1$, we can Taylor expand the Bessel functions in \eqref{eq:ops_time} about $t=0$ to get
 \begin{align}
-  \bk{\Psi|\O\p{t}|\Psi}
-  \approx 1 - \f12 \p{tJ}^2 \sum_{\mu,\nu}
+  \bk{\Psi|s_{\mu\nu}\p{t}|\Psi}
+  &\approx \bk{\psi\op{\mu}{\nu}\psi}
+  \times \sp{1 - \f14 \p{t J}^2 \sin\p{\xi_{\mu\nu}^-}^2}, \\
+  \bk{\Psi|s^2\p{t}|\Psi}
+  &\approx 1 - \f12 \p{tJ}^2 \sum_{\mu,\nu}
   \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2 \sin\p{\xi_{\mu\nu}^-}^2.
 \end{align}
-In the limit of a small angle $\xi$ for which $I\xi\ll1$, we can
-further simplify
+In the limit of a small angle $\xi$ for which $I\xi\ll1$, we can further simplify
 \begin{align}
-  \bk{\Psi|\O\p{t}|\Psi}
-  \approx 1 - \f14 \var_\psi\p{Z} \p{JI\xi}^2 \times t^2,
+  \bk{\Psi|s_{\mu\nu}\p{t}|\Psi}
+  &\approx \bk{\psi\op{\mu}{\nu}\psi}
+  \times \sp{1 - \f1{16} \p{\f{\mu-\nu}{I}}^2
+    \p{JI\xi}^2 \times t^2}, \\
+  \bk{\Psi|s^2\p{t}|\Psi}
+  &\approx 1 - \f14 \var_\psi\p{Z} \p{JI\xi}^2 \times t^2,
   \label{eq:small_time_angle}
@@
 \end{align}
-Note that the result in \eqref{eq:small_time_angle} is valid for all
-$t\ll\p{JI\xi}^{-1}$, which may be a long time given the small-angle
-assumption $I\xi\ll1$.  Figure \ref{fig:limiting_vals} shows several
-values of $\var_\psi\p{Z}$ for different $n$.
+Note that the result in \eqref{eq:small_time_angle} is valid for all $t\ll\p{JI\xi}^{-1}$, which may be a long time given the small-angle assumption $I\xi\ll1$.
+Figure \ref{fig:limiting_vals} shows several values of $\var_\psi\p{Z}$ for different $n$.
 
@@
 
-We now consider the opposite, strongly-interacting limit of a 1D
-model,
+We now consider the strongly-interacting limit of a 1D model, in which the Hamiltonian is
+\begin{align}
+  H = -\f{U}{N} \v S^\dag\cdot \v S
+  - \f{J}{2} \sum_{\mu,q} \cos\p{q + \mu\xi} s_{\mu\mu q}
+\end{align}
+with $J\ll U$.
 

### 2020-08-12 sun_phases/sun_dynamics_notes.tex
change some notation for compatibility with other notes start writing explanation of nuclear spin tomography

@@
 \begin{align}
-  H = -\f{J}{2} \sum_{\mu,q} \cos\p{q + \mu\xi} s_{\mu\mu q},
+  H = -\f{J}{2} \sum_{\mu,q} \cos\p{q + \mu\varphi} s_{\mu\mu q},
 \end{align}
-where $q$ is a quasi-momentum (in units with lattice spacing $a=1$); $\mu\in\set{I,I-1,\cdots,-I}$ is a nuclear spin projection onto a quantization axis, with $I\equiv\p{n-1}/2$ the total nuclear spin; $\xi$ is a spin-orbit-coupling angle, and $J$ is the single-particle bandwidth (proportional to the tunneling rate, $J/4$).
+where $q$ is a quasi-momentum (in units with lattice spacing $a=1$); $\mu\in\set{I,I-1,\cdots,-I}$ is a nuclear spin projection onto a quantization axis, with $I\equiv\p{n-1}/2$ the total nuclear spin; $\varphi$ is a spin-orbit-coupling angle, and $J$ is the single-particle bandwidth (proportional to the tunneling rate, $J/4$).
 As dynamical order parameters, we consider the operators
@@
   \eta_{\mu\nu q}
-  \equiv -\f12 \sp{\cos\p{q+\mu\xi} - \cos\p{q+\nu\xi}}
-  = \sin\p{\xi_{\mu\nu}^-} \sin\p{p+\xi_{\mu\nu}^+},
+  \equiv -\f12 \sp{\cos\p{q+\mu\varphi} - \cos\p{q+\nu\varphi}}
+  = \sin\p{\varphi_{\mu\nu}^-} \sin\p{p+\varphi_{\mu\nu}^+},
   &&
-  \xi_{\mu\nu}^\pm \equiv \f{\mu\pm\nu}{2}\,\xi.
+  \varphi_{\mu\nu}^\pm \equiv \f{\mu\pm\nu}{2}\,\varphi.
 \end{align}
@@
   = \f1N \bk{\psi\op{\mu}{\nu}\psi} \sum_q e^{\i t J \eta_{\mu\nu q}}
-  = \bk{\psi\op{\mu}{\nu}\psi} \chi_{\mu\nu}\p{\xi},
+  = \bk{\psi\op{\mu}{\nu}\psi} \chi_{\mu\nu}\p{\varphi},
 \end{align}
@@
   &= \sum_{\mu,\nu} \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2
-  \abs{\chi_{\mu\nu}\p{\xi}}^2 + \f{n-1}{N},
+  \abs{\chi_{\mu\nu}\p{\varphi}}^2 + \f{n-1}{N},
 \end{align}
@@
 \begin{align}
-  \chi_{\mu\nu}\p{\xi} \equiv \f1N \sum_q e^{\i tJ\eta_{\mu\nu q}}
-  = \f1N \sum_q e^{\i tJ\sin\p{\xi_{\mu\nu}^-} \sin\p{q+\xi_{\mu\nu}^+}}.
+  \chi_{\mu\nu}\p{\varphi} \equiv \f1N \sum_q e^{\i tJ\eta_{\mu\nu q}}
+  = \f1N \sum_q e^{\i tJ\sin\p{\varphi_{\mu\nu}^-} \sin\p{q+\varphi_{\mu\nu}^+}}.
 \end{align}
@@
 \begin{align}
-  \chi_{\mu\nu}\p{\xi} \approx \f1{2\pi} \int_{-\pi}^\pi \dd\theta\,
-  e^{\i tJ\sin\p{\xi_{\mu\nu}^-} \sin\p{\theta+\xi_{\mu\nu}^+}}
-  = \J_0\sp{tJ\sin\p{\xi_{\mu\nu}^-}},
+  \chi_{\mu\nu}\p{\varphi} \approx \f1{2\pi} \int_{-\pi}^\pi \dd\theta\,
+  e^{\i tJ\sin\p{\varphi_{\mu\nu}^-} \sin\p{\theta+\varphi_{\mu\nu}^+}}
+  = \J_0\sp{tJ\sin\p{\varphi_{\mu\nu}^-}},
 \end{align}
@@
   &\approx \bk{\psi\op{\mu}{\nu}\psi}
-  \J_0\sp{tJ\sin\p{\xi_{\mu\nu}^-}},
+  \J_0\sp{tJ\sin\p{\varphi_{\mu\nu}^-}},
   \\
@@
   &\approx \sum_{\mu,\nu} \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2
-  \J_0\sp{tJ\sin\p{\xi_{\mu\nu}^-}}^2.
+  \J_0\sp{tJ\sin\p{\varphi_{\mu\nu}^-}}^2.
   \label{eq:ops_time}
@@
 
-The constant (time independent) terms in \eqref{eq:ops_time} have $\mu=\nu$ (in which case all $\xi_{\mu\nu}^-=0$), so up to $O(1/N)$ corrections
+The constant (time independent) terms in \eqref{eq:ops_time} have $\mu=\nu$ (in which case all $\varphi_{\mu\nu}^-=0$), so up to $\O(1/N)$ corrections
 \begin{align}
   \bk{\Psi|s_{\mu\nu}^{\t{const}}|\Psi}
-  \approx \abs{\bk{\mu|\psi}}^2,
+  \approx \delta_{\mu\nu} \abs{\bk{\mu|\psi}}^2,
   &&
   \bk{\Psi|s^2_{\t{const}}|\Psi}
-  \approx r_\psi \equiv \sum_\mu \abs{\bk{\mu|\psi}}^4,
+  \approx \sum_\mu \abs{\bk{\mu|\psi}}^4 \equiv r_\psi,
   \label{eq:ops_const}
@@
 
-At ``short'' times $t$ for which $tJ\sin\p{I\xi}\ll1$, we can Taylor expand the Bessel functions in \eqref{eq:ops_time} about $t=0$ to get
+At ``short'' times $t$ for which $tJ\sin\p{I\varphi}\ll1$, we can Taylor expand the Bessel functions in \eqref{eq:ops_time} about $t=0$ to get
 \begin{align}
@@
   &\approx \bk{\psi\op{\mu}{\nu}\psi}
-  \times \sp{1 - \f14 \p{t J}^2 \sin\p{\xi_{\mu\nu}^-}^2}, \\
+  \times \sp{1 - \f14 \p{t J}^2 \sin\p{\varphi_{\mu\nu}^-}^2}, \\
   \bk{\Psi|s^2\p{t}|\Psi}
   &\approx 1 - \f12 \p{tJ}^2 \sum_{\mu,\nu}
-  \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2 \sin\p{\xi_{\mu\nu}^-}^2.
+  \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2 \sin\p{\varphi_{\mu\nu}^-}^2.
 \end{align}
-In the limit of a small angle $\xi$ for which $I\xi\ll1$, we can further simplify
+In the limit of a small angle $\varphi$ for which $I\varphi\ll1$, we can further simplify
 \begin{align}
@@
   \times \sp{1 - \f1{16} \p{\f{\mu-\nu}{I}}^2
-    \p{JI\xi}^2 \times t^2}, \\
+    \p{JI\varphi}^2 \times t^2}, \\
   \bk{\Psi|s^2\p{t}|\Psi}
-  &\approx 1 - \f14 \var_\psi\p{Z} \p{JI\xi}^2 \times t^2,
+  &\approx 1 - \f14 \var_\psi\p{Z} \p{JI\varphi}^2 \times t^2,
   \label{eq:small_time_angle}
@@
 \end{align}
-Note that the result in \eqref{eq:small_time_angle} is valid for all $t\ll\p{JI\xi}^{-1}$, which may be a long time given the small-angle assumption $I\xi\ll1$.
+Note that the result in \eqref{eq:small_time_angle} is valid for all $t\ll\p{JI\varphi}^{-1}$, which may be a long time given the small-angle assumption $I\varphi\ll1$.
 Figure \ref{fig:limiting_vals} shows several values of $\var_\psi\p{Z}$ for different $n$.
@@
   H = -\f{U}{N} \v S^\dag\cdot \v S
-  - \f{J}{2} \sum_{\mu,q} \cos\p{q + \mu\xi} s_{\mu\mu q}
+  - \f{J}{2} \sum_{\mu,q} \cos\p{q + \mu\varphi} s_{\mu\mu q}
 \end{align}
 with $J\ll U$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newpage
+\section{Nuclear spin tomography}
+
+Here we discuss the task of performing full tomography on the average reduced single-particle density operator of a collection of nuclear spins.
+As this task will be performed using collective measurements and homogeneous control fields, for ease of language we will consider the equivalent problem of performing tomography on a single nuclear spin.
+We have essentially two ingredients at our disposal: projective measurements of spin onto a fixed quantization axis, and a three-laser drive that addresses nuclear spins via off-resonant coupling to an excited electronic state.
+To be concrete, we can directly measure projectors $\op{\mu}$ onto states of definite spin projection $\mu$ onto the $z$ axis.
+While the three-laser drive gives us access to a variety of nuclear spin Hamiltonians, in notably it allows us to implement arbitrary SU(2) rotations of the form $e^{-\i\v\theta\c\v S}$, with $\v S\equiv\p{S_\x,S_\y,S_\z}$ a vector of spin operators and $\v\theta\equiv\p{\theta_\x,\theta_\y,\theta_\z}$ an arbitrary rotation vector.
+These rotations essentially allow us to measure projectors $\Pi_{\v v\mu} \equiv \op{\mu_{\v v}}$ onto states of definite spin projection $\mu$ onto an arbitrary quantization axis $\v v$.
+
+Question: given a collection of projectors $\Pi\equiv\set{\Pi_j}$ and measurement outcomes $M\equiv\set{M_j}$, with $M_j$ an empirical estimate of $\bk{\Pi_j}_\rho=\tr\p{\rho\Pi_j}$, how do we determine whether the measurement data $M$ is sufficient to reconstruct $\rho$?
+Answer: reconstruction of $\rho$ is possible if $\Pi$ spans the entire space $\B_n$ of operators on the Hilbert space $\H_n$ of an $n$-level spin.
+Given an orthonormal basis $\set{Q_\alpha}$ of self-adjoint operators spanning $\B_n$, we can expand
+\begin{align}
+  \rho = \sum_\alpha \rho_\alpha Q_\alpha,
+  &&
+  \rho_\alpha \equiv \tr\p{\rho Q_\alpha}.
+\end{align}
+If $\Pi$ spans $\B_n$, then we can find a set of real numbers $c_{\alpha j}\in\RR$ for which
+\begin{align}
+  Q_\alpha = \sum_j c_{\alpha j} \Pi_j.
+\end{align}
+An empirical estimate $\tilde\rho$ of $\rho$ is then
+\begin{align}
+  \tilde\rho = \sum_\alpha \tilde\rho_\alpha Q_\alpha,
+  &&
+  \tilde\rho_\alpha \equiv \sum_j c_{\alpha j} M_j.
+\end{align}
+This estimate can be improved by defining $\tilde\rho$ as the least-squares fit to the set of linear equations
+\begin{align}
+  \tr\p{\tilde\rho\,\Pi_j} = M_j,
+\end{align}
+and performing maximum-likelihood corrections to $\tilde\rho$ \cite{smolin2012efficient}, which essentially consists of
+\begin{enumerate*}
+\item diagonalizing $\tilde\rho$,
+\item eliminating the most negative eigenvalue (setting it to zero),
+\item adding an equal amount to all other eigenvalues to enforce $\tr\tilde\rho=1$, and
+\item repeating steps (ii,iii) until there are no more negative eigenvalues.
+\end{enumerate*}
+This algorithm finds the closest density operator to $\tilde\rho$ that has no negative eigenvalues, where distance is measured by the metric induced by the operator 2-norm: $\norm{\O}_2\equiv\sqrt{\tr\p{\O^\dag \O}}$.
+
+As it turns out, numerical experiments with $n\le20$ suggest that in order to perfurm full tomography of $\rho$, it suffices to measure all projectors $\Pi_{\v v\mu}$ onto states of definite spin ($\mu$) along $2n-1$ different axes ($\v v$).
+The question remains: which axes $\set{\v v}$ should we choose?
+
+\bibliography{sun_dynamics_notes.bib}
 

### 2020-08-13 sun_phases/sun_dynamics_notes.tex
added discussion for finding a 'good' set of measurement axes for nuclear spin tomography

@@
 To be concrete, we can directly measure projectors $\op{\mu}$ onto states of definite spin projection $\mu$ onto the $z$ axis.
-While the three-laser drive gives us access to a variety of nuclear spin Hamiltonians, in notably it allows us to implement arbitrary SU(2) rotations of the form $e^{-\i\v\theta\c\v S}$, with $\v S\equiv\p{S_\x,S_\y,S_\z}$ a vector of spin operators and $\v\theta\equiv\p{\theta_\x,\theta_\y,\theta_\z}$ an arbitrary rotation vector.
-These rotations essentially allow us to measure projectors $\Pi_{\v v\mu} \equiv \op{\mu_{\v v}}$ onto states of definite spin projection $\mu$ onto an arbitrary quantization axis $\v v$.
+While the three-laser drive gives us access to a variety of nuclear spin Hamiltonians, most notably it allows us to implement arbitrary SU(2) rotations of the form $e^{-\i\v\theta\c\v S}$, with $\v S\equiv\p{S_\x,S_\y,S_\z}$ a vector of spin operators and $\v\theta\equiv\p{\theta_\x,\theta_\y,\theta_\z}$ an arbitrary rotation vector.
+These rotations essentially allow us to measure projectors $\Pi_{\v v\mu} \equiv \op{\mu_{\v v}}$ onto states of definite spin projection $\mu$ onto an arbitrary quantization axis $\v v\in\SS_2/\ZZ_2$.
 
 Question: given a collection of projectors $\Pi\equiv\set{\Pi_j}$ and measurement outcomes $M\equiv\set{M_j}$, with $M_j$ an empirical estimate of $\bk{\Pi_j}_\rho=\tr\p{\rho\Pi_j}$, how do we determine whether the measurement data $M$ is sufficient to reconstruct $\rho$?
-Answer: reconstruction of $\rho$ is possible if $\Pi$ spans the entire space $\B_n$ of operators on the Hilbert space $\H_n$ of an $n$-level spin.
+Answer: reconstruction of $\rho$ is possible if $\Pi$ spans the entire space $\B_n$ of operators on the Hilbert space of an $n$-level spin.
 Given an orthonormal basis $\set{Q_\alpha}$ of self-adjoint operators spanning $\B_n$, we can expand
@@
   &&
-  \rho_\alpha \equiv \tr\p{\rho Q_\alpha}.
+  \rho_\alpha \equiv \bk{Q_\alpha}_\rho.
 \end{align}
@@
 \end{align}
-An empirical estimate $\tilde\rho$ of $\rho$ is then
-\begin{align}
-  \tilde\rho = \sum_\alpha \tilde\rho_\alpha Q_\alpha,
-  &&
-  \tilde\rho_\alpha \equiv \sum_j c_{\alpha j} M_j.
-\end{align}
-This estimate can be improved by defining $\tilde\rho$ as the least-squares fit to the set of linear equations
-\begin{align}
-  \tr\p{\tilde\rho\,\Pi_j} = M_j,
-\end{align}
-and performing maximum-likelihood corrections to $\tilde\rho$ \cite{smolin2012efficient}, which essentially consists of
+An empirical estimate $\tilde\rho$ of $\rho$ is then\footnote{This estimate can be improved by defining $\tilde\rho$ as the least-squares fit to the set of linear equations $\tr\p{\tilde\rho\,\Pi_j} = M_j$, and performing maximum-likelihood corrections to $\tilde\rho$ \cite{smolin2012efficient}, which essentially consists of
 \begin{enumerate*}
@@
 \end{enumerate*}
-This algorithm finds the closest density operator to $\tilde\rho$ that has no negative eigenvalues, where distance is measured by the metric induced by the operator 2-norm: $\norm{\O}_2\equiv\sqrt{\tr\p{\O^\dag \O}}$.
-
-As it turns out, numerical experiments with $n\le20$ suggest that in order to perfurm full tomography of $\rho$, it suffices to measure all projectors $\Pi_{\v v\mu}$ onto states of definite spin ($\mu$) along $2n-1$ different axes ($\v v$).
-The question remains: which axes $\set{\v v}$ should we choose?
+This algorithm finds the closest density operator to $\tilde\rho$ that has no negative eigenvalues, where distance is measured by the metric induced by the operator 2-norm: $\norm{\O}_2\equiv\sqrt{\tr\p{\O^\dag \O}}$.}
+\begin{align}
+  \tilde\rho = \sum_\alpha \tilde\rho_\alpha Q_\alpha,
+  &&
+  \tilde\rho_\alpha \equiv \sum_j c_{\alpha j} M_j
+  \approx \sum_j c_{\alpha j} \bk{\Pi_j}_\rho
+  = \bk{Q_\alpha}_\rho
+  = \rho_\alpha.
+\end{align}
+As it turns out, numerical experiments with $n\le20$ suggest that in order to perfurm full tomography of $\rho$, it suffices to measure all projectors $\Pi_{\v v\mu}$ onto states of definite spin $\mu$ along $2n-1$ different axes $\v v$.
+The question remains: which axes should we choose?
+As a first pass, we can simply choose a set of $2n-1$ random axes, $V\equiv\set{\v v}$, by uniformly sampling points on the sphere.
+These random axes define a set of $\p{2n-1}\times n$ projectors $\Pi_V=\set{\Pi^V_j}$, where for shorthand we use a combined index $j=\p{\v v_j,\mu_j}$ that specifies both a measurement axis $\v v_j$ and an outcome $\mu_j$, and for clarity we index objects by $V$ to emphasize their dependence on the choice of measurement axes.
+The projectors $\Pi_V$ can be used to construct $n^2$ linearly independent operators by constructing a matrix $B_V$ with entries
+\begin{align}
+  \sp{B_V}_{ij} = \tr\p{\Pi^V_i \Pi^V_j}.
+\end{align}
+Indexing normalized eigenvectors and the corresponding eigenvalues of $B_V$ by $\alpha$, respectively $\v x^V_\alpha$ and $m^V_\alpha$, for all $\alpha$ with eigenvalue $m^V_\alpha\ne0$ we can then construct the normalized and linearly independent operators
+\begin{align}
+  Q^V_\alpha \equiv \f1{\sqrt{m^V_\alpha}} \sum_j x^V_{\alpha j} \Pi^V_j.
+\end{align}
+To summarize: randomly choosing a set $V$ of $2n-1$ axes will induce a set of measurement operators $\Pi_V$ from which we can construct a complete basis $\set{Q^V_\alpha}$ of operators for $\B_n$.
+However, some choices of axes will be better than others: a good choice of axes $V$ will maximize the amount of information that the corresponding measurement data $M_V=\set{M^V_j}$ will contain about the expectation values $\bk{Q^V_\alpha}_\rho$ used to reconstruct $\rho$.
+The amount of information that $M_V$ contains about $\bk{Q^V_\alpha}_\rho$ is controlled, in part, by the magnitude of the eigenvalue $m^V_\alpha$: if $m^V_\alpha$ is large, then $M_V$ will generally contain a lot of information about $\bk{Q^V_\alpha}_\rho$, and vice versa\footnote{Think about the limit $m^V_\alpha\to0$, in which case the fact that $Q^V_\alpha$ is normalized implies that the operator $\sum_j x^V_{\alpha j} \Pi^V_j$ must have a vanishing norm.
+  This vanishing norm can only occur due to cancellations that are analogous to a ``sign problem,'' making the expectation value of $\sum_j x^V_{\alpha j} \Pi^V_j$ difficult to estimate because it will be dominated by measurement noise.}.
+Rather than choosing axes randomly, we can therefore choose a set of axes that minimizes some cost function $C\p{m_V}$ favoring large eigenvalues $m_V\equiv\set{m^V_\alpha}$.
+A reasonable cost function that treats all eigenvalues on equal footing, for example, is
+\begin{align}
+  C_{\t{inv}}\p{m_V} \equiv \sum_\alpha \f1{m^V_\alpha},
+\end{align}
+where the sum is implicitly performed over the $n^2$ largest eigenvalues of $B_V$ (as all other eigenvalues are guaranteed to be zero).
+Note that any information about $\rho$, obtained either from prior knowledge or preliminary measurement data, can be used to construct an even better choice of measurement axes, leading to the possibility of adaptive measurement protocols \cite{pereira2018adaptive} that are more efficient in terms of the number of measurements required to estimate $\rho$ to a fixed precision.
+We leave the details of tailored and adaptive measurement protocols to future work.
 

### 2020-08-18 sun_phases/multilevel_spin_notes.tex
add some drive hamiltonians
add more drive Hamiltonians
mionr fixes

@@
 
-Rather than coupling the states of a multilevel system directly with a
-single driving laser, as in Section \ref{sec:drive_SOC}, we now
-consider coupling through an off-resonant Raman transition to a highly
-excited state.  Specifically, we consider a three-laser drive of the
-form
+Rather than coupling the states of a multilevel system directly with a single driving laser, as in Section \ref{sec:drive_SOC}, we now consider coupling through an off-resonant Raman transition to a highly excited state.
+Specifically, we consider a three-laser drive of the form
 \begin{align}
@@
 \end{align}
-where $j$ indexes a lattice site, $\Omega_m$ is a (real) driving
-amplitude; $\phi_m$ is the relative phase of drive $m$ between
-adjacent lattice sites; $\op{r}{s}$ for $r,s\in\set{\up,\dn}$ is a
-pseudo-spin operator for the auxiliary degree of freedom that is
-off-resonantly addressed by the lasers; $\Delta$ is a detuning of the
-lasers from the energy of an excited state; and $\1$ is the identity
-operator.  If the detuning is large, $\abs{\Delta}\gg\abs{\Omega_m}$,
-then at second order in perturbation theory the effective action of
-$H_{\t{Raman,bare}}^{(\phi)}$ within the $\ket{\dn}$-state manifold
-takes the form
+where $j$ indexes a lattice site, $\Omega_m$ is a (real) driving amplitude; $\phi_m$ is the relative phase of drive $m$ between adjacent lattice sites; $\op{r}{s}$ for $r,s\in\set{\up,\dn}$ is a pseudo-spin operator for the auxiliary degree of freedom that is off-resonantly addressed by the lasers; $\Delta$ is a detuning of the lasers from the energy of an excited state; and $\1$ is the identity operator.
+If the detuning is large, $\abs{\Delta}\gg\abs{\Omega_m}$, then at second order in perturbation theory the effective action of $H_{\t{Raman,bare}}^{(\phi)}$ within the $\ket{\dn}$-state manifold takes the form
 \begin{align}
@@
 \end{align}
-where $H_{\t{Raman},\tau}^{(\phi)}$ generates nuclear spin transitions
-$\mu\to\mu\pm\tau$:
+where $H_{\t{Raman},\tau}^{(\phi)}$ generates nuclear spin transitions $\mu\to\mu\pm\tau$:
 \begin{align}
@@
 \end{align}
-where the coefficients $\Omega_{LM}^{(\phi,j)}$ are provided in
-Appendix \ref{sec:drive_raman_coeff}.  If the driving lasers are
-phase-matched with $\phi_m=m\phi$, such that a nuclear spin transition
-$\mu\to\mu+m$ on site $j$ imprints the phase $m\phi j$, then the drive
-$H_{\t{Raman}}^{(\phi)}$ takes the form
+where the coefficients $\Omega_{LM}^{(\phi,j)}$ are provided in Appendix \ref{sec:drive_raman_coeff}.
+If the driving lasers are phase-matched with $\phi_m=m\phi$, such that a nuclear spin transition $\mu\to\mu+m$ on site $j$ imprints the phase $m\phi j$, then the drive $H_{\t{Raman}}^{(\phi)}$ takes the form
 \begin{align}
@@
 \end{align}
-with rotated collective drive operators $\D_{LM}^{(\phi)}$ defined in
-\eqref{eq:drive_rot} and uniform driving amplitudes $\Omega_{LM}$
-provided in Appendix \ref{sec:drive_raman_coeff}.  Similarly to the
-case of a direct drive considered in Section \ref{sec:drive_SOC}, the
-Raman drive $H_{\t{Raman}}^{(\phi)}$ in \eqref{eq:drive_raman_matched}
-can be made homogenous through the gauge transformation
-$c_{j\mu}^\dag \to e^{-i\mu\phi j} c_{j\mu}^\dag$, which takes
+with rotated collective drive operators $\D_{LM}^{(\phi)}$ defined in \eqref{eq:drive_rot} and uniform driving amplitudes $\Omega_{LM}$ provided in Appendix \ref{sec:drive_raman_coeff}.
+Similarly to the case of a direct drive considered in Section \ref{sec:drive_SOC}, the Raman drive $H_{\t{Raman}}^{(\phi)}$ in \eqref{eq:drive_raman_matched} can be made homogenous through the gauge transformation $c_{j\mu}^\dag \to e^{-i\mu\phi j} c_{j\mu}^\dag$, which takes
 \begin{align}
@@
 \end{align}
-Writing $H_{\t{Raman}} = \sum_j H_{\t{Raman}}^{\t{single},j}$, where
-$H_{\t{Raman}}^{\t{single},j}$ denotes the action of
-$H_{\t{Raman}}^{\t{single}}$ on spin $j$, we can expand
+Writing $H_{\t{Raman}} = \sum_j H_{\t{Raman}}^{\t{single},j}$, where $H_{\t{Raman}}^{\t{single},j}$ denotes the action of $H_{\t{Raman}}^{\t{single}}$ on spin $j$, we can expand
 \begin{align}
@@
 \end{align}
-where
-$\p{S_\z,S_\x,S_\y} \equiv \xi_1\times\p{D_{1,0},D_{1,1},D_{1,-1}}$ is
-the standard SU(2) subalgebra of the operators on a single $n$-level
-spin, and
+where $\p{S_\z,S_\x,S_\y} \equiv \xi_1\times\p{D_{1,0},D_{1,1},D_{1,-1}}$ is the standard $\mathfrak{su}(2)$ subalgebra of the operators on a single $n$-level spin, and
 \begin{align}
@@
 \end{align}
-Tuning the amplitudes $\Omega_0,\Omega_+,\Omega_-$, or equivalently
-$h_0,h_+,h_-$, enables us to implement various drive Hamiltonians,
-several of which are shown in Table \ref{tab:drives}.
+Tuning the amplitudes $\Omega_0,\Omega_+,\Omega_-$, or equivalently $h_0,h_+,h_-$, enables us to implement various drive Hamiltonians, several of which are shown in Table \ref{tab:drives}.
 
@@
   \centering
-  \caption{Drive Hamiltonians that can be implemented with various
-    amplitude-matching conditions.}
+  \caption{Drive Hamiltonians that can be implemented with various amplitude-matching conditions.}
   \begin{tabular}{c|c|c}
     Condition & $\p{h_0,h_+,h_-}/h$
-    & $H_{\t{Raman}}^{\t{single}}/\p{h^2/\Delta}$ \\ \hline\hline
-    $\Omega_0=\Omega_+=+\Omega_-$ & $\p{1,\sqrt{2},0}$
-    & $\sqrt{2} S_\x + S_\x^2 - S_\y^2$ \\ \hline
-    $\Omega_0=\Omega_+=-\Omega_-$ & $\p{1,0,\sqrt{2}}$
-    & $\sqrt{2}\p{S_\z S_\x + S_\x S_\z} - \p{S_\x^2-S_\y^2}$ \\ \hline
-    $\Omega_+=\Omega_-=0$ & $\p{1,0,0}$
-    & $S_\x^2 + S_\y^2$ \\ \hline
-    $\Omega_0=\Omega_-=0$ & $\p{0,1,+1}$
-    & $+S_\z - \p{S_\x^2 + S_\y^2}$ \\ \hline
-    $\Omega_0=\Omega_+=0$ & $\p{0,1,-1}$
-    & $-S_\z - \p{S_\x^2 + S_\y^2}$
+    & $H_{\t{Raman}}^{\t{single}}/\p{h^2/\Delta}$
+    \\ \hline\hline
+    $\Omega_0=\pm\Omega_+=\pm\Omega_-$ & $\p{1,\pm\sqrt{2},0}$
+    & $\pm\sqrt{2} S_\x + \p{S_\x^2 - S_\y^2}$
+    \\ \hline
+    $\Omega_0=\pm\Omega_+=\mp\Omega_-$ & $\p{1,0,\pm\sqrt{2}}$
+    & $\pm\sqrt{2}\p{S_\z S_\x + S_\x S_\z} - \p{S_\x^2-S_\y^2}$
+    \\ \hline
+    $\Omega_0=\Omega_\mp=0$ & $\p{0,1,\pm1}$
+    & $\pm S_\z + S_\z^2$
+    \\ \hline
+    $\sqrt{2}\Omega_0=\pm\Omega_+=\pm\Omega_-$
+    & $\p{1,\pm1,0}$ & $\pm S_\x + S_\x^2$
+    \\ \hline
+    $\sqrt{2}\Omega_0=\pm\Omega_+=\mp\Omega_-$
+    & $\p{1,0,\pm1}$ & $\pm\p{S_\z S_\x+S_\x S_\z} + S_\y^2$
+    \\ \hline
+    $\Omega_+=\Omega_-=0$ & $\p{1,0,0}$ & $-S_\z^2$
+    \\ \hline
+    $\Omega_0=0$,  $\Omega_+=+\Omega_-$ & $\p{0,1,0}$ & $-S_\y^2$
+    \\ \hline
+    $\Omega_0=0$,  $\Omega_+=-\Omega_-$ & $\p{0,0,1}$ & $-S_\x^2$
   \end{tabular}

### 2020-08-18 sun_phases/sun_dynamics_notes.tex
mionr fixes

@@
 \begin{align}
-  H = \f1N \sum_{\substack{\mu,\nu,\rho,\sigma\\i<j}}
+  H = \f1N \sum_{i<j}
   h_{ij} \Pi_{ij}
@@
   &&
-  \Pi_{ij} \equiv \sum_{\mu,\nu} s_{\mu\nu i} s_{\rho\sigma j}.
+  \Pi_{ij} \equiv \sum_{\mu,\nu} s_{\mu\nu i} s_{\nu\mu j}.
   \label{eq:spin_sun}
@@
   s^2 \equiv \v s^\dag\cdot \v s
-  = \f1{N^2} \sum_{\mu,\nu,p,q} s_{\mu\nu p} s_{\nu\mu q},
+  = \f1{N^2} \sum_{\mu,\nu,p,q} s_{\mu\nu p} s_{\nu\mu q}.
 \end{align}
-where $\v s_q\equiv\sum_{\mu,\nu} s_{\mu\nu q}\uv e_{\mu\nu}$ is a vector of normalized generators of $U(n)$ for spin $q$, and $\uv e_{\mu\nu}$ is a unit vector of $\RR^{n^2}$.
 In the Heisenberg picture, these operators at time $t$ are
@@
   \equiv -\f12 \sp{\cos\p{q+\mu\varphi} - \cos\p{q+\nu\varphi}}
-  = \sin\p{\varphi_{\mu\nu}^-} \sin\p{p+\varphi_{\mu\nu}^+},
+  = \sin\p{\varphi_{\mu\nu}^-} \sin\p{q+\varphi_{\mu\nu}^+},
   &&
@@
 \end{align}
-An empirical estimate $\tilde\rho$ of $\rho$ is then\footnote{This estimate can be improved by defining $\tilde\rho$ as the least-squares fit to the set of linear equations $\tr\p{\tilde\rho\,\Pi_j} = M_j$, and performing maximum-likelihood corrections to $\tilde\rho$ \cite{smolin2012efficient}, which essentially consists of
-\begin{enumerate*}
-\item diagonalizing $\tilde\rho$,
-\item eliminating the most negative eigenvalue (setting it to zero),
-\item adding an equal amount to all other eigenvalues to enforce $\tr\tilde\rho=1$, and
-\item repeating steps (ii,iii) until there are no more negative eigenvalues.
-\end{enumerate*}
-This algorithm finds the closest density operator to $\tilde\rho$ that has no negative eigenvalues, where distance is measured by the metric induced by the operator 2-norm: $\norm{\O}_2\equiv\sqrt{\tr\p{\O^\dag \O}}$.}
+An empirical estimate $\tilde\rho$ of $\rho$ is then\footnote{This estimate can be improved by defining $\tilde\rho$ as the least-squares fit to the set of linear equations $\tr\p{\tilde\rho\,\Pi_j} = M_j$, and performing maximum-likelihood corrections to $\tilde\rho$ \cite{smolin2012efficient}.}
 \begin{align}
@@
 \end{align}
-Indexing normalized eigenvectors and the corresponding eigenvalues of $B_V$ by $\alpha$, respectively $\v x^V_\alpha$ and $m^V_\alpha$, for all $\alpha$ with eigenvalue $m^V_\alpha\ne0$ we can then construct the normalized and linearly independent operators
+Indexing the (normalized) eigenvectors and the corresponding eigenvalues of $B_V$ by $\alpha$, respectively $\v x^V_\alpha$ and $m^V_\alpha$, for all $\alpha$ with eigenvalue $m^V_\alpha\ne0$ we can then construct the normalized and linearly independent operators
 \begin{align}
@@
 However, some choices of axes will be better than others: a good choice of axes $V$ will maximize the amount of information that the corresponding measurement data $M_V=\set{M^V_j}$ will contain about the expectation values $\bk{Q^V_\alpha}_\rho$ used to reconstruct $\rho$.
-The amount of information that $M_V$ contains about $\bk{Q^V_\alpha}_\rho$ is controlled, in part, by the magnitude of the eigenvalue $m^V_\alpha$: if $m^V_\alpha$ is large, then $M_V$ will generally contain a lot of information about $\bk{Q^V_\alpha}_\rho$, and vice versa\footnote{Think about the limit $m^V_\alpha\to0$, in which case the fact that $Q^V_\alpha$ is normalized implies that the operator $\sum_j x^V_{\alpha j} \Pi^V_j$ must have a vanishing norm.
-  This vanishing norm can only occur due to cancellations that are analogous to a ``sign problem,'' making the expectation value of $\sum_j x^V_{\alpha j} \Pi^V_j$ difficult to estimate because it will be dominated by measurement noise.}.
+The amount of information that $M_V$ contains about $\bk{Q^V_\alpha}_\rho$ is determined, in part, by the magnitude of the eigenvalue $m^V_\alpha$: if $m^V_\alpha$ is small, then $M_V$ will generally contain little information about $\bk{Q^V_\alpha}_\rho$, and vice versa\footnote{Think about the limit $m^V_\alpha\to0$, in which case the fact that $Q^V_\alpha$ is normalized implies that the operator $\sum_j x^V_{\alpha j} \Pi^V_j$ must have a vanishing norm.
+  A vanishing norm can only occur due to fine-tuned cancellations that make the expectation value of $\sum_j x^V_{\alpha j} \Pi^V_j$ difficult to estimate due to the presence of measurement errors and shot noise.}.
 Rather than choosing axes randomly, we can therefore choose a set of axes that minimizes some cost function $C\p{m_V}$ favoring large eigenvalues $m_V\equiv\set{m^V_\alpha}$.

### 2020-08-18 collective_sun/scripts/plot_squeezed_gaussian.py

@@ -29,5 +29,10 @@ plt.yticks([])
 plt.xlabel("$X$")
 plt.ylabel("$P$")
 
-plt.tight_layout(pad = 0)
+plt.arrow(+max_val*.8, 0, -max_val/4, 0, color = "c", head_width = .2)
+plt.arrow(-max_val*.8, 0, +max_val/4, 0, color = "c", head_width = .2)
+plt.arrow(0, +max_val*.45, 0, +max_val/4, color = "c", head_width = .2)
+plt.arrow(0, -max_val*.45, 0, -max_val/4, color = "c", head_width = .2)
+
+plt.tight_layout(pad = 0.1)
 plt.savefig(fig_dir + "squeezed_gaussian.pdf")

### 2020-08-19 sun_phases/multilevel_spin_notes.tex
minor fix

@@
 \end{align}
-where $\p{S_\z,S_\x,S_\y} \equiv \xi_1\times\p{D_{1,0},D_{1,1},D_{1,-1}}$ is the standard $\mathfrak{su}(2)$ subalgebra of the operators on a single $n$-level spin, and
+where $\v S = \p{S_\z,S_\x,S_\y} \equiv \xi_1\times\p{D_{1,0},D_{1,1},D_{1,-1}}$ is the standard $\mathfrak{su}(2)$ subalgebra of the operators on a single $n$-level spin, and
 \begin{align}
@@
     $\Omega_0=\Omega_\mp=0$ & $\p{0,1,\pm1}$
-    & $\pm S_\z + S_\z^2$
+    & $\pm S_\z + S_\z^2 - \v S^2$
     \\ \hline
@@
     \\ \hline
-    $\Omega_+=\Omega_-=0$ & $\p{1,0,0}$ & $-S_\z^2$
+    $\Omega_+=\Omega_-=0$ & $\p{1,0,0}$ & $-S_\z^2 + \v S^2$
     \\ \hline

### 2020-08-25 sun_phases/sun_dynamics_notes.tex
added initial state with back-to-back spins

@@
 The amount of information that $M_V$ contains about $\bk{Q^V_\alpha}_\rho$ is determined, in part, by the magnitude of the eigenvalue $m^V_\alpha$: if $m^V_\alpha$ is small, then $M_V$ will generally contain little information about $\bk{Q^V_\alpha}_\rho$, and vice versa\footnote{Think about the limit $m^V_\alpha\to0$, in which case the fact that $Q^V_\alpha$ is normalized implies that the operator $\sum_j x^V_{\alpha j} \Pi^V_j$ must have a vanishing norm.
-  A vanishing norm can only occur due to fine-tuned cancellations that make the expectation value of $\sum_j x^V_{\alpha j} \Pi^V_j$ difficult to estimate due to the presence of measurement errors and shot noise.}.
+  A vanishing norm can only occur due to fine-tuned cancellations that make the expectation value of $\sum_j x^V_{\alpha j} \Pi^V_j$ difficult to estimate from noisy data on $\bk{\Pi^V_j}_\rho$.}.
 Rather than choosing axes randomly, we can therefore choose a set of axes that minimizes some cost function $C\p{m_V}$ favoring large eigenvalues $m_V\equiv\set{m^V_\alpha}$.

### 2020-08-25 sun_phases/scripts/plot_bosons.py

@@ -19,8 +19,7 @@ else:
     init_state_str = "X"
 
 assert( spin_num % 2 == 0 )
-assert( init_state_str in [ "X", "DS" ] )
-if spin_dim == 2: init_state_str = "X"
+assert( init_state_str in [ "X", "XX", "DS" ] )
 
 figsize = (4,3)
 figsize_ext = (5,3)

### 2020-08-25 sun_phases/scripts/simulate_bosons.py

@@ -16,8 +16,7 @@ else:
     init_state_str = "X"
 
 assert( spin_num % 2 == 0 )
-assert( init_state_str in [ "X", "DS" ] )
-if spin_dim == 2: init_state_str = "X"
+assert( init_state_str in [ "X", "XX", "DCS" ] )
 
 # simulation parameters
 log10_tun_vals = np.linspace(-2,1,13)
@@ -130,9 +129,14 @@ def evolve(initial_state, sim_time, field, coupling_op = None,
     initial_state.shape = state_shape
     return times, states
 
+# get the angle associated with a given spin number <--> quasimomentum
+def spin_angle(qq):
+    return 2*np.pi*qq/spin_num
+
 # construct an on-site external field
 def bare_lattice_field(qq, soc_angle):
-    return np.diag([ np.cos(2*np.pi*qq/spin_num + mu * soc_angle) for mu in spin_vals ])
+    return np.diag([ np.cos(spin_angle(qq) + mu * soc_angle) for mu in spin_vals ])
+
 
 ##########################################################################################
 # set up objects for simulation
@@ -150,7 +154,15 @@ def compute_spin_mats(states):
 # construct initial state
 if init_state_str == "X":
     init_state = polarized_state([0,1,0])
-elif init_state_str == "DS":
+elif init_state_str == "XX":
+    # polarized along +X (-X) for low- (high-)momentum spins
+    spin_states = [ spin_state([0,+1,0]) if np.cos(spin_angle(qq+1/spin_num)) > 0 else
+                    spin_state([0,-1,0])
+                    for qq in range(spin_num) ]
+    init_state = boson_mft_state(spin_states)
+elif init_state_str == "DCS":
+    # "double coherent state", with every spin is in an (identical) superposition
+    #   of two coherent states
     alpha = np.pi/2 + np.arcsin(1/3)
     beta = np.pi/3
     gamma = -(spin_dim-1)/3 * np.pi
@@ -161,6 +173,7 @@ elif init_state_str == "DS":
 ##########################################################################################
 # simulate!
 ##########################################################################################
+soc_frac_vals = soc_frac_vals[soc_frac_vals != 0]
 
 sim_start = time.time()
 

### 2020-09-08 sun_phases/sun_dynamics_notes.tex
change convention with factor of 2, add some text about interactions

@@
 \newcommand{\X}{\text{X}}
-\newcommand{\DS}{\text{DS}}
+\newcommand{\DCS}{\text{DCS}}
 
@@
 \begin{align}
-  \ket{\DS}
+  \ket{\DCS}
   \equiv \f1{\sqrt{2}}
@@
 \begin{align}
-  H = -\f{J}{2} \sum_{\mu,q} \cos\p{q + \mu\varphi} s_{\mu\mu q},
+  H = -J \sum_{\mu,q} \cos\p{q + \mu\varphi} s_{\mu\mu q},
 \end{align}
-where $q$ is a quasi-momentum (in units with lattice spacing $a=1$); $\mu\in\set{I,I-1,\cdots,-I}$ is a nuclear spin projection onto a quantization axis, with $I\equiv\p{n-1}/2$ the total nuclear spin; $\varphi$ is a spin-orbit-coupling angle, and $J$ is the single-particle bandwidth (proportional to the tunneling rate, $J/4$).
+where $q$ is a quasi-momentum (in units with lattice spacing $a=1$); $\mu\in\set{I,I-1,\cdots,-I}$ is a nuclear spin projection onto a quantization axis, with $I\equiv\p{n-1}/2$ the total nuclear spin; $\varphi$ is a spin-orbit-coupling angle; and $J$ is essentially the single-particle bandwidth (proportional to the nearest-neighbor tunneling rate, $J/2$).
 As dynamical order parameters, we consider the operators
@@
   s_{\mu\nu}\p{t} = \f1N \sum_q s_{\mu\nu q}
-  \times e^{\i t J \eta_{\mu\nu q}},
+  \times e^{\i\eta_{\mu\nu q} J t},
   &&
   s^2\p{t} = \f1{N^2} \sum_{\mu,\nu,p,q} s_{\mu\nu p} s_{\nu\mu q}
-  \times e^{\i tJ\p{\eta_{\mu\nu p}-\eta_{\mu\nu q}}},
+  \times e^{\i \p{\eta_{\mu\nu p}-\eta_{\mu\nu q}} J t},
 \end{align}
@@
   \eta_{\mu\nu q}
-  \equiv -\f12 \sp{\cos\p{q+\mu\varphi} - \cos\p{q+\nu\varphi}}
-  = \sin\p{\varphi_{\mu\nu}^-} \sin\p{q+\varphi_{\mu\nu}^+},
+  \equiv -\sp{\cos\p{q+\mu\varphi} - \cos\p{q+\nu\varphi}}
+  = 2 \sin\p{\varphi_{\mu\nu}^-} \sin\p{q+\varphi_{\mu\nu}^+},
   &&
@@
   \bk{\Psi|s_{\mu\nu}\p{t}|\Psi}
-  = \f1N \bk{\psi\op{\mu}{\nu}\psi} \sum_q e^{\i t J \eta_{\mu\nu q}}
+  = \f1N \bk{\psi\op{\mu}{\nu}\psi} \sum_q e^{\i\eta_{\mu\nu q} J t}
   = \bk{\psi\op{\mu}{\nu}\psi} \chi_{\mu\nu}\p{\varphi},
@@
   &= \f1{N^2} \sum_{\mu,\nu} \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2
-  \sum_{p\ne q} e^{\i tJ\p{\eta_{\mu\nu p}-\eta_{\mu\nu q}}} + \f{n}{N} \\
+  \sum_{p\ne q} e^{\i\p{\eta_{\mu\nu p}-\eta_{\mu\nu q}} J t} + \f{n}{N} \\
   &= \sum_{\mu,\nu} \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2
@@
 \begin{align}
-  \chi_{\mu\nu}\p{\varphi} \equiv \f1N \sum_q e^{\i tJ\eta_{\mu\nu q}}
-  = \f1N \sum_q e^{\i tJ\sin\p{\varphi_{\mu\nu}^-} \sin\p{q+\varphi_{\mu\nu}^+}}.
+  \chi_{\mu\nu}\p{\varphi} \equiv \f1N \sum_q e^{\i\eta_{\mu\nu q} J t}
+  = \f1N \sum_q e^{\i 2\sin\p{\varphi_{\mu\nu}^-} \sin\p{q+\varphi_{\mu\nu}^+} J t}.
 \end{align}
@@
   \chi_{\mu\nu}\p{\varphi} \approx \f1{2\pi} \int_{-\pi}^\pi \dd\theta\,
-  e^{\i tJ\sin\p{\varphi_{\mu\nu}^-} \sin\p{\theta+\varphi_{\mu\nu}^+}}
-  = \J_0\sp{tJ\sin\p{\varphi_{\mu\nu}^-}},
+  e^{\i 2\sin\p{\varphi_{\mu\nu}^-} \sin\p{\theta+\varphi_{\mu\nu}^+} J t}
+  = \J_0\sp{2 \sin\p{\varphi_{\mu\nu}^-} J t},
 \end{align}
@@
   &\approx \bk{\psi\op{\mu}{\nu}\psi}
-  \J_0\sp{tJ\sin\p{\varphi_{\mu\nu}^-}},
+  \J_0\sp{2\sin\p{\varphi_{\mu\nu}^-}Jt},
   \\
@@
   &\approx \sum_{\mu,\nu} \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2
-  \J_0\sp{tJ\sin\p{\varphi_{\mu\nu}^-}}^2.
+  \J_0\sp{2\sin\p{\varphi_{\mu\nu}^-}Jt}^2.
   \label{eq:ops_time}
@@
 
-At ``short'' times $t$ for which $tJ\sin\p{I\varphi}\ll1$, we can Taylor expand the Bessel functions in \eqref{eq:ops_time} about $t=0$ to get
+At ``short'' times $t$ for which $\sin\p{I\varphi}Jt\ll1$, we can Taylor expand the Bessel functions in \eqref{eq:ops_time} about $t=0$ to get
 \begin{align}
@@
   &\approx \bk{\psi\op{\mu}{\nu}\psi}
-  \times \sp{1 - \f14 \p{t J}^2 \sin\p{\varphi_{\mu\nu}^-}^2}, \\
+  \times \sp{1 - \sin\p{\varphi_{\mu\nu}^-}^2 \p{J t}^2}, \\
   \bk{\Psi|s^2\p{t}|\Psi}
-  &\approx 1 - \f12 \p{tJ}^2 \sum_{\mu,\nu}
-  \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2 \sin\p{\varphi_{\mu\nu}^-}^2.
+  &\approx 1 - 2 \p{Jt}^2 \sum_{\mu,\nu}
+  \sin\p{\varphi_{\mu\nu}^-}^2 \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2.
 \end{align}
@@
   &\approx \bk{\psi\op{\mu}{\nu}\psi}
-  \times \sp{1 - \f1{16} \p{\f{\mu-\nu}{I}}^2
-    \p{JI\varphi}^2 \times t^2}, \\
+  \times \sp{1 - \p{\f{\mu-\nu}{2I}}^2
+    \p{I\varphi J t}^2}, \\
   \bk{\Psi|s^2\p{t}|\Psi}
-  &\approx 1 - \f14 \var_\psi\p{Z} \p{JI\varphi}^2 \times t^2,
+  &\approx 1 - \var_\psi\p{Z} \p{I\varphi J t}^2,
   \label{eq:small_time_angle}
@@
 \end{align}
-Note that the result in \eqref{eq:small_time_angle} is valid for all $t\ll\p{JI\varphi}^{-1}$, which may be a long time given the small-angle assumption $I\varphi\ll1$.
+Note that the result in \eqref{eq:small_time_angle} is valid for all $t\ll\p{I\varphi J}^{-1}$, which may be a long time given the small-angle assumption $I\varphi\ll1$.
 Figure \ref{fig:limiting_vals} shows several values of $\var_\psi\p{Z}$ for different $n$.
@@
     different $n$.  Computed with respect to the states
-    $\ket\X\equiv\ket{\pi/2,0}$ and $\ket{\DS}$, defined in
+    $\ket\X\equiv\ket{\pi/2,0}$ and $\ket{\DCS}$, defined in
     \eqref{eq:polarized_state} and \eqref{eq:double_state}.}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Strongly-interacting limit of a realistic 1D model}
+\section{Interacting regime of a realistic 1D model}
 
-We now consider the strongly-interacting limit of a 1D model, in which the Hamiltonian is
+We now consider a realistic 1D model with interactions, in which the Hamiltonian is
 \begin{align}
   H = -\f{U}{N} \v S^\dag\cdot \v S
-  - \f{J}{2} \sum_{\mu,q} \cos\p{q + \mu\varphi} s_{\mu\mu q}
+  - J \sum_{\mu,q} \cos\p{q + \mu\varphi} s_{\mu\mu q}.
 \end{align}
-with $J\ll U$.
+This Hamiltonian can be written in the simplified form
+\begin{align}
+  H = -\f{U}{N} \v S^\dag\cdot \v S
+  - J \sum_q \p{\cos q\, w_{\varphi,q}^+ - \sin q\, w_{\varphi,q}^-},
+\end{align}
+where the diagonal single-spin operators
+\begin{align}
+  w_\varphi^+ \equiv \sum_\mu \cos\p{\mu\varphi} \op{\mu},
+  &&
+  w_\varphi^- \equiv \sum_\mu \sin\p{\mu\varphi} \op{\mu},
+\end{align}
+are respectively even ($+$) and odd ($-$) under spin inversion, $\mu\to-\mu$.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
 \section{Nuclear spin tomography}

### 2020-09-14 sun_phases/notes.tex
split off sun_dynamics_notes.tex

--- 2020-09-08 sun_phases/sun_dynamics_notes.tex
@@
 While the three-laser drive gives us access to a variety of nuclear spin Hamiltonians, most notably it allows us to implement arbitrary SU(2) rotations of the form $e^{-\i\v\theta\c\v S}$, with $\v S\equiv\p{S_\x,S_\y,S_\z}$ a vector of spin operators and $\v\theta\equiv\p{\theta_\x,\theta_\y,\theta_\z}$ an arbitrary rotation vector.
-These rotations essentially allow us to measure projectors $\Pi_{\v v\mu} \equiv \op{\mu_{\v v}}$ onto states of definite spin projection $\mu$ onto an arbitrary quantization axis $\v v\in\SS_2/\ZZ_2$.
+These rotations essentially allow us to measure projectors $\Pi_{\v v\mu} \equiv \op{\mu_{\v v}}$ onto states of definite spin projection $\mu$ onto an arbitrary quantization axis $\v v\in\SS_2$.
 
@@
 \end{align}
-To summarize: randomly choosing a set $V$ of $2n-1$ axes will induce a set of measurement operators $\Pi_V$ from which we can construct a complete basis $\set{Q^V_\alpha}$ of operators for $\B_n$.
+To summarize: randomly choosing a set $V$ of $2n-1$ axes will determine a set of measurement operators $\Pi_V$ from which we can construct a complete basis $\set{Q^V_\alpha}$ of operators for $\B_n$.
 However, some choices of axes will be better than others: a good choice of axes $V$ will maximize the amount of information that the corresponding measurement data $M_V=\set{M^V_j}$ will contain about the expectation values $\bk{Q^V_\alpha}_\rho$ used to reconstruct $\rho$.
@@
 
-\bibliography{sun_dynamics_notes.bib}
+\bibliography{notes.bib}
 

### 2020-09-14 sun_phases/sun_dynamics_notes.tex
split off sun_dynamics_notes.tex

--- 2020-08-25 sun_phases/sun_dynamics_notes.tex
@@
 
+\usepackage{setspace} % to change text spacing
+
 %%% symbols, notations, etc.
@@
 \newcommand{\uv}[1]{\bm{\hat{#1}}} % unit vectors
-\newcommand{\av}{\vec} % arrow vectors
 \renewcommand{\c}{\cdot} % inner product
@@
 
+\renewcommand{\a}{\text{a}}
 \newcommand{\x}{\text{x}}
@@
 \newcommand{\X}{\text{X}}
-\newcommand{\DS}{\text{DS}}
+\newcommand{\Y}{\text{Y}}
+\newcommand{\Z}{\text{Z}}
+\newcommand{\DCS}{\text{DCS}}
 
 \newcommand{\B}{\mathcal{B}}
-\newcommand{\D}{\mathcal{D}}
-\newcommand{\E}{\mathcal{E}}
-\renewcommand{\H}{\mathcal{H}}
-\newcommand{\I}{\mathcal{I}}
-\newcommand{\J}{\mathcal{J}}
-\newcommand{\M}{\mathcal{M}}
-\newcommand{\N}{\mathcal{N}}
-\renewcommand{\O}{\mathcal{O}}
-\renewcommand{\P}{\mathcal{P}}
-\newcommand{\Q}{\mathcal{Q}}
-\newcommand{\R}{\mathcal{R}}
-\newcommand{\T}{\mathcal{T}}
-\renewcommand{\S}{\mathcal{S}}
-\newcommand{\V}{\mathcal{V}}
-\newcommand{\Z}{\mathcal{Z}}
 
-\newcommand{\EE}{\mathbb{E}}
 \newcommand{\RR}{\mathbb{R}}
@@
 \newcommand{\ZZ}{\mathbb{Z}}
-
-\newcommand{\PS}{\text{PS}}
-\newcommand{\col}{\underline}
-
-\let\var\relax
-\DeclareMathOperator{\var}{var}
 
@@
 \newcommand{\red}[1]{{\color{red} #1}}
+\newcommand{\todo}[1]{{\color{magenta} To do: #1}}
 
@@
 
-% todo: find roots of lax vector
-
-\title{Collective SU($n$) spin model dynamics}%
+\title{Collective SU($n$) spin dynamics}%
 \author{Michael A. Perlin}%
@@
 
+% notes for Emil Yuzbashyan
+
 \maketitle
+\begin{singlespace}
+  \tableofcontents
+\end{singlespace}
+\vspace{1cm}
 
-We first consider a general quadratic spin Hamiltonian of the form
+Multilevel atoms on an optical lattice can be used to implement a variety of fermionic and spin models with SU($n$) symmetries.
+Not only are these systems powerful quantum simulators, they are also promising platforms for quantum information processing and metrology.
+Two-level atoms with SU(2)-symmetric interactions, for example, can be used to simulate dynamical phases of BCS superconductors, or to prepare entangled states that enable surpassing classical limits on measurement precision.
+When atoms have $n$ nuclear spin sublevels, inter-atomic interactions can exhibit an SU($n$) symmetry that has yet to be explored as a scientific resource.
+For this reason, we wish to study the dynamical quantum phases that can be implemented with these atoms.
+
+Here, we will consider a particular collective spin model resembling that used to study BCS superconductivity.
+We begin by briefly deriving the spin model, before discussing some of its symmetries, examples of initial states that can be prepared, and additional experimental controls.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{The spin model}
+
+Starting from a Hamiltonian describing multilevel atoms on a 1D lattice, we first derive the spin model that we will consider throughout the remainder of these notes, in \eqref{eq:spin_cos} and \eqref{eq:spin_pair}.
+Everything before \eqref{eq:spin_cos} is provided for context, but the details are (at this stage) not important for the spin physics that we will discuss.
+
+Our starting point is a simple 1D model of $n$-level fermions with nearest-neighbor tunneling and on-site (collisional) exchange interactions:
 \begin{align}
-  H = \f1N \sum_{\substack{\mu,\nu,\rho,\sigma\\i<j}}
-  h^{\mu\nu i}_{\rho\sigma j} s_{\mu\nu i} s_{\rho\sigma j}
-  + \sum_{\mu,\nu,i} \epsilon_{\mu\nu i} s_{\mu\nu i},
-  \label{eq:spin}
+  H_{\t{bare}} = -\f{J}{2} \sum_{j,\mu}
+  \p{c_{j\mu}^\dag c_{j+1,\mu} + \t{h.c.}}
+  + U \sum_{j,\mu,\nu} c_{j\mu}^\dag c_{j\mu} c_{j\nu}^\dag c_{j\nu},
+  \label{eq:bare_sites}
 \end{align}
-where $\mu,\nu\in\set{I,I-1,\cdots,-I}$ with $I\equiv\p{n-1}/2$ index orthogonal states of an $n$-level spin; $i,j\in\ZZ_N$ index one of $N$ spins; $h^{\mu\nu i}_{\rho\sigma j}$ and $\epsilon_{\mu\nu i}$ are scalars; and $s_{\mu\nu i}\coloneqq\op{\mu}{\nu}_i$ is a transition operator for spin $i$.
-If interactions are SU($n$) symmetric then $h^{\mu\nu i}_{\rho\sigma k} = h_{ik} \delta_{\mu\sigma}\delta_{\nu\rho}$, in which case
+where $c_{j\mu}$ annihilates a fermionic atom on lattice site $j$ with nuclear spin projection $\mu\in\set{I,I-1,\cdots,-I}$ onto a quantization axis; $I\equiv\p{n-1}/2$ is the total nuclear spin of each atom (e.g.~$I=9/2$ in the case of Sr-87 with $n=10$ nuclear spin levels); $J$ is a tunneling rate; and $U$ is an on-site interaction strength.
+In terms of quasi-momentum modes, the same Hamiltonian reads
 \begin{align}
-  H = \f1N \sum_{i<j}
-  h_{ij} \Pi_{ij}
-  + \sum_{\mu,\nu,i} \epsilon_{\mu\nu i} s_{\mu\nu i},
+  H_{\t{bare}} = -J \sum_{q,\mu} \cos q\, c_{q\mu}^\dag c_{q\mu}
+  + \f{U}{L} \sum_{\substack{k,\ell,p,q\\\mu,\nu}}
+  \delta_{k+\ell,p+q}\, c_{k\mu}^\dag c_{p\mu} c_{\ell\nu}^\dag c_{q\nu},
+  \label{eq:bare_momenta}
+\end{align}
+where $L$ is the number of lattice sites; $k,\ell,p,q$ index quasi-momenta; and $\delta_{k+\ell,p+q}$ enforces momentum conservation.
+If interactions are sufficiently weak, $U\lesssim J$, then we can neglect mode-changing collisions and only keep interaction terms in which the final momenta are the same as the initial momenta, $\set{p,q}=\set{k,\ell}$.
+In this case, the ``direct'' terms with $\p{p,q}=\p{k,\ell}$ turn out to be trivial (proportional to the total atom number squared, $N^2$), leaving only ``exchange'' terms with $\p{p,q}=\p{\ell,k}$.
+In total, the operator content of the interactions can be written in the form
+\begin{align}
+  \sum_{\substack{k,\ell,p,q\\\mu,\nu}}
+  \delta_{k+\ell,p+q}\, c_{k\mu}^\dag c_{p\mu} c_{\ell\nu}^\dag c_{q\nu}
+  \to \sum_{k,\ell,\mu,\nu} c_{k\mu}^\dag c_{\ell\mu}
+  c_{\ell\nu}^\dag c_{k\nu}
+  = -\sum_{k,\ell,\mu,\nu} s_{\mu\nu k} s_{\nu\mu\ell}
+  = -\sum_{k,\ell} \v s_k \c\v s_\ell
+  = -\v S \c\v S,
+\end{align}
+where we define spin operators $s_{\mu\nu k}\equiv c_{\mu k}^\dag c_{\nu k}$; $\v s_k=\sum_{\mu,\nu}s_{\mu\nu k}\op{\mu}{\nu}$ is an operator-valued spin matrix for mode $k$; $\v S\equiv\sum_k \v s_k$ is a collective spin matrix; and $\v A\c\v B\equiv\tr\p{\v A^\dag \v B}$ is an inner product of spin matrices (whose result is, again, operator-valued).
+Putting everything together, we arrive at a spin Hamiltonian of the form
+\begin{align}
+  H_{\t{spin}}^{\t{bare}}
+  = -\f{U}{L} \v S \c\v S - J \sum_{q,\mu} \cos q\, s_{\mu\mu q}.
+  \label{eq:bare_spin}
+\end{align}
+The Hamiltonian we have thus derived governs free evolution, without external control fields.
+External lasers can additionally couple atoms' internal states, but at the cost of momentum kicks that generally induce spin-orbit coupling (see Appendix \ref{sec:controls}).
+To diagonalize these laser drives in their orbital (lattice site or quasi-momentum) index, it turns out to be convenient to make the gauge transformation $c_{j\mu}\to e^{\i\mu\phi j} c_{j\mu}$ for some spin-orbit coupling angle $\phi$, after which the Hamiltonian in \eqref{eq:bare_spin} becomes
+\begin{align}
+  H_{\t{spin}} = -\f{U}{L} \v S \c\v S
+  - J \sum_{q,\mu} \cos\p{q+\mu\phi}\, s_{\mu\mu q}.
+  \label{eq:spin_cos}
+\end{align}
+The resulting dispersion relation is shown in Figure \ref{fig:dispersion}.
+This Hamiltonian can be written in the simplified form
+\begin{align}
+  H_{\t{spin}} = -\f{U}{L} \v S\c\v S
+  - J \sum_q \p{\cos q\, w_{\phi,q}^+ - \sin q\, w_{\phi,q}^-},
+  \label{eq:spin_pair}
+\end{align}
+where the diagonal single-spin operators
+\begin{align}
+  w_\phi^+ \equiv \sum_\mu \cos\p{\mu\phi} s_{\mu\mu},
   &&
-  \Pi_{ij} \equiv \sum_{\mu,\nu} s_{\mu\nu i} s_{\nu\mu j}.
-  \label{eq:spin_sun}
+  w_\phi^- \equiv \sum_\mu \sin\p{\mu\phi} s_{\mu\mu},
+\end{align}
+are respectively even ($+$) and odd ($-$) under spin inversion, $\mu\to-\mu$.
+
+\begin{figure}
+  \centering
+  \includegraphics{figures/sun_dispersion.pdf}
+  \caption{Single-particle dispersion relations $E_{q\mu}=-J\cos\p{q+\mu\phi}$ for the case of $n=6$ (nuclear spin $I=5/2$) and different spin-orbit coupling angles $\phi$.
+    Legend indicates the spin projection $\mu$ onto the $z$.
+    Note that these dispersion relations are ``maximally spread'' when $\phi=2\pi/n$ (as with $\phi=\pi/3$ above), and that degeneracies occur at all $q$ when $\phi=2\pi/m$ for any $m\in\set{1,2,\cdots,n-1}$.}
+  \label{fig:dispersion}
+\end{figure}
+
+In the case of SU(2), $w_\phi^+=\cos\p{\phi/2}\1$ is a scalar and $w_\phi^-=\sin\p{\phi/2}\sigma_\z$, so
+\begin{align}
+  H_{\t{spin}}^{(n=2)} = -\f{U}{L} \v S\c\v S
+  + J \sin\p{\phi/2} \sum_q \sin q\, \sigma_{\z,q}
+  - J \cos\p{\phi/2} \sum_q \cos q\, \1_q,
+\end{align}
+where we include the identity terms for comparison with \eqref{eq:spin_pair}.
+If $n$ is arbitrary but $I\phi\ll1$, then similarly
+\begin{align}
+  H_{\t{spin}} = -\f{U}{L} \v S\c\v S
+  + J \phi \sum_q \sin q\, s_{\z,q}
+  - J \sum_q \cos q\, \1_q + O(\phi^2),
+  &&
+  s_\z \equiv \sum_\mu \mu\, s_{\mu\mu}.
+\end{align}
+Finally, if $\phi=\pi$ then $\cos\p{\mu\pi}=0$ and $\sin\p{\mu\pi}=\p{-1}^{\mu-1/2}$ due the fact that $\mu$ is a half-integer\footnote{Bosonic nuclei with $I\ne0$ are unsable, so all $I\ne0$ and $\mu\in\set{I,I-1,\cdots,-I}$ are half-integer-valued.}, so
+\begin{align}
+  H_{\t{spin}}^{(\phi=\pi)} = -\f{U}{L} \v S\c\v S
+  + J \sum_q \sin q\, s_{\a,q},
+  &&
+  s_\a \equiv \sum_\mu \p{-1}^{\mu-1/2} s_{\mu\mu}.
+  \label{eq:spin_pi}
 \end{align}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Schwinger boson representation}
+\section{Conserved quantities and collective observables}
 
-We can write the Hamiltonian in \eqref{eq:spin} using Schwinger bosons as
-\begin{align}
-  H = \f1N \sum_{\substack{\mu,\nu,\rho,\sigma\\i<j}}
-  h^{\mu\nu i}_{\rho\sigma j}
-  b_{\mu i}^\dag b_{\nu i} b_{\rho j}^\dag b_{\sigma j}
-  + \sum_{\mu,\nu,i} \epsilon_{\mu\nu i} b_{\mu i}^\dag b_{\nu i},
-\end{align}
-where $b_{\mu i}$ annihilates a boson of type $\mu$ on site $i$.
-The Heisenberg equations of motion for these operators are\footnote{The Hamiltonian in \eqref{eq:spin} only defines $h^{\mu\nu j}_{\rho\sigma k}$ for $j<k$.
-  To simplify expressions in the remainder of this work, we therefore additionally define $h^{\mu\nu k}_{\rho\sigma j}=h^{\rho\sigma j}_{\mu\nu k}$ and $h^{\mu\nu i}_{\rho\sigma i}=0$.}
-\begin{align}
-  \i \d_t b_{\alpha i} = \sp{b_{\alpha i}, H}
-  &= \f1N \sum_{\substack{\mu,\nu,\rho,\sigma\\j<k}}
-  h^{\mu\nu j}_{\rho\sigma k}
-  \sp{b_{\alpha i}, b_{\mu j}^\dag b_{\nu j} b_{\rho k}^\dag b_{\sigma k}}
-  + \sum_{\mu,\nu,j} \epsilon_{\mu\nu j}
-  \sp{b_{\alpha i}, b_{\mu j}^\dag b_{\nu j}} \\
-  &= \f1N \sum_{\mu,\nu,\rho,\sigma,k}
-  h^{\mu\nu j}_{\rho\sigma k}
-  \sp{b_{\alpha i}, b_{\mu i}^\dag b_{\nu i}} b_{\rho k}^\dag b_{\sigma k}
-  + \sum_{\mu,\nu} \epsilon_{\mu\nu i}
-  \sp{b_{\alpha i}, b_{\mu i}^\dag b_{\nu i}} \\
-  &= \sum_{\mu,\nu} \p{\f1N \sum_{\rho,\sigma,k}
-    h^{\mu\nu i}_{\rho\sigma k} b_{\rho k}^\dag b_{\sigma k}
-    + \epsilon_{\mu\nu i}}
-  \sp{b_{\alpha i}, b_{\mu i}^\dag b_{\nu i}}
-\end{align}
-where
-\begin{align}
-  \sp{b_{\alpha i}, b_{\mu i}^\dag b_{\nu i}}
-  = \delta_{\mu\alpha} \delta_{\nu\alpha} b_{\alpha i}
-  + \delta_{\mu\alpha} \p{1-\delta_{\nu\alpha}} b_{\nu i}
-  = \delta_{\mu\alpha} b_{\nu i},
-\end{align}
-so
-\begin{align}
-  \i \d_t b_{\alpha i}
-  = \sum_\nu \p{\f1N \sum_{\rho,\sigma,k}
-    h^{\alpha\nu i}_{\rho\sigma k} b_{\rho k}^\dag b_{\sigma k}
-    + \epsilon_{\alpha\nu i}} b_{\nu i}.
-\end{align}
+Needless to say, the single-body terms in \eqref{eq:spin_pair} explicitly break the SU($n$) symmetry of the interactions.
+Even so, the fact that all single-body terms are diagonal with respect to nuclear spin projection $\mu$ implies that the overall Hamiltonian conserves all collective spin operators of the form $S_{\mu\mu}\equiv\sum_q s_{\mu\mu q}$, as well as all products thereof, which is equivalent to conserving all powers of $S_\z$ in the case of SU(2).
+The enormous number of these conserved quantities\footnote{Altogether, there are essentially ${N+n-1 \choose n-1}\sim N^{n-1}$ such quantities.} suggests that the spin Hamiltonian in \eqref{eq:spin_pair} is integrable, although the precise form of an appropriate Bethe ansatz or Lax formulation is unclear.
+The situation is somewhat simpler when $\phi\ll1$ or $\phi=\pi$, in which case the single-body terms are built from a single spin operator (namely, $s_\z$ or $s_\a$).
+
+It is worth mentioning at this point that we can, in principle, perform full tomography on the average single-body reduced density operator $\bar\rho$, which is equivalent to measuring all components of the mean collective spin matrix $\bk{\v S}\equiv\sum_{\mu,\nu}\bk{S_{\mu\nu}}\op{\mu}{\nu}$ (see Appendix \ref{sec:tomography} for additional details).
+The squared ``magnetization'' $\bk{\v S}\c\bk{\v S}=\sum_{\mu,\nu}\abs{\bk{S_{\mu\nu}}}^2$, which is essentially equal to the purity of $\bar\rho$, may therefore be an interesting ``unbiased'' dynamical order parameter to consider.
+In addition, diagonalizing $\bar\rho$ (equivalently, $\bk{\v S}$), provides a ``natural'' basis in which to examine collective observables.
+The eigenvectors and eigenvalues of $\bar\rho$ may therefore be used as additional order parameters for identifying distinct dynamical phases.
+The precise interpretation of these order parameters is worth ironing out and discussing in the future.
 
@@
 
-We can essentially prepare $N$-fold tensor products $\ket\Psi\equiv\ket\psi^{\otimes N}$ of three types of initial product states $\ket\psi$.
-First and foremost, we can prepare states $\ket{\mu}$ of definite spin projection $\mu=\set{I,I-1,\cdots,-I}$ onto a quantization axis.
-In addition, we can prepare arbitrary polarized states defined by the polar and azimuthal angles $\theta,\phi$:
-\begin{align}
-  \ket{\theta,\phi}
-  \equiv e^{i\phi s_\z} e^{i\theta s_\y} \ket\up
-  = \sum_\mu \zeta_\mu\p{\theta} e^{-\i\mu\phi} \ket\mu,
-  \label{eq:polarized_state}
-\end{align}
-where $\ket\up\equiv\ket{I}$ is a state polarized along the $z$ axis, and\footnote{The expansion of $\ket{\theta,\phi}$ in \eqref{eq:polarized_state} and \eqref{eq:polarized_proj} is acquired by representing an $n$-level spin by $n-1$ qubits restricted to the (Dicke) manifold of permutationally symmetric states.}
-\begin{align}
-  \zeta_\mu\p{\theta} \equiv { 2I \choose I+\mu }^{1/2}
-  \cos\p{\f{\theta}{2}}^{I+\mu} \sin\p{\f{\theta}{2}}^{I-\mu}.
-  \label{eq:polarized_proj}
-\end{align}
-Finally, for even $n$ we can also prepare the following superposition of polarized states:
-\begin{align}
-  \ket{\DS}
-  \equiv \f1{\sqrt{2}}
-  \p{e^{\i\gamma} \ket{\alpha,\beta}
-    + e^{-\i\gamma} \ket{\alpha,-\beta}}
-  = \sum_\mu \zeta_\mu\p{\alpha}
-  \times \sqrt{2}\, \cos\p{\mu\beta-\gamma} \ket\mu
-  \label{eq:double_state}
-\end{align}
-with
-\begin{align}
-  \alpha = \f{\pi}{2} + \arcsin\p{\f13},
-  &&
-  \beta = \f{\pi}{3},
-  &&
-  \gamma = -\f{2\pi}{3}\, I.
-\end{align}
+The simplest initial state we can prepare is that with all atoms occupying a single nuclear spin state, e.g.~the state $\ket{\Z}\equiv\ket{I}^{\otimes N}$ with all $N$ atoms pointing up along the $z$ axis.
+Using external control fields (discussed in the Section \ref{sec:controls}), we can rotate the state $\ket\Z$ to orient it along any axis in 3D space.
+External control fields also allow us to prepare states analogous to $\p{\ket\up+\ket\dn}^{\otimes N}$ in that each spin is in a coherent superposition of states polarized along different directions, e.g.~$\p{\ket{I}+\ket{-I}}^{\otimes N}$.
+The possibilities to prepare $N$-fold product states of the form $\ket\psi^{\otimes N}$ are enormous.
+
+Perhaps more interestingly, we can in principle prepare inhomogeneous product states in which the state of each spin is correlated with its corresponding single-body term of the spin Hamiltonian in \eqref{eq:spin_pair}.
+We are currently working on characterizing the sort of inhomogeneous states that we can prepare.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Non-interacting limit of a realistic 1D model}
+\section{Mean-field theory}
 
-We consider the single-particle Hamiltonian
+The dynamics induced by the spin Hamiltonian in \eqref{eq:spin_cos} and \eqref{eq:spin_pair} are efficiently simulable when $U=0$ or $U\gg JI\phi$.
+In order to explore the entirety of parameter space, however, we turn to mean-field theory.
+To this end, we first represent individual spin operators by Schwinger bosons, $s_{\mu\nu q}\coloneqq b_{\mu q}^\dag b_{\nu q}$.
+Up to an overall constant, our spin model can then be written in the form
 \begin{align}
-  H = -\f{J}{2} \sum_{\mu,q} \cos\p{q + \mu\varphi} s_{\mu\mu q},
+  H_{\t{boson}} = - \f{U}{L} \sum_{\substack{\mu,\nu\\p\ne q}}
+  b_{\mu p}^\dag b_{\nu p} b_{\nu q}^\dag b_{\mu q}
+  - J \sum_{\mu,q} \cos\p{q+\mu\phi} b_{\mu q}^\dag b_{\mu q}.
 \end{align}
-where $q$ is a quasi-momentum (in units with lattice spacing $a=1$); $\mu\in\set{I,I-1,\cdots,-I}$ is a nuclear spin projection onto a quantization axis, with $I\equiv\p{n-1}/2$ the total nuclear spin; $\varphi$ is a spin-orbit-coupling angle, and $J$ is the single-particle bandwidth (proportional to the tunneling rate, $J/4$).
-As dynamical order parameters, we consider the operators
+With a bit of algebra, we can work out that the Heisenberg equations of motion for individual boson operators are
 \begin{align}
-  s_{\mu\nu} \equiv \f1N \sum_q s_{\mu\nu q},
-  &&
-  s^2 \equiv \v s^\dag\cdot \v s
-  = \f1{N^2} \sum_{\mu,\nu,p,q} s_{\mu\nu p} s_{\nu\mu q}.
+  \i \d_t b_{\mu q} = \sp{b_{\mu q},H}
+  = -2\, \f{U}{L} \sum_{\nu,p} b_{\nu p}^\dag b_{\mu p} b_{\nu q}
+  - J \cos\p{q+\mu\phi} b_{\mu q}.
 \end{align}
-In the Heisenberg picture, these operators at time $t$ are
-\begin{align}
-  s_{\mu\nu}\p{t} = \f1N \sum_q s_{\mu\nu q}
-  \times e^{\i t J \eta_{\mu\nu q}},
-  &&
-  s^2\p{t} = \f1{N^2} \sum_{\mu,\nu,p,q} s_{\mu\nu p} s_{\nu\mu q}
-  \times e^{\i tJ\p{\eta_{\mu\nu p}-\eta_{\mu\nu q}}},
-\end{align}
-where
-\begin{align}
-  \eta_{\mu\nu q}
-  \equiv -\f12 \sp{\cos\p{q+\mu\varphi} - \cos\p{q+\nu\varphi}}
-  = \sin\p{\varphi_{\mu\nu}^-} \sin\p{q+\varphi_{\mu\nu}^+},
-  &&
-  \varphi_{\mu\nu}^\pm \equiv \f{\mu\pm\nu}{2}\,\varphi.
-\end{align}
-In particular, we are interested in the expectation values
-\begin{align}
-  \bk{\Psi|s_{\mu\nu}\p{t}|\Psi}
-  = \f1N \bk{\psi\op{\mu}{\nu}\psi} \sum_q e^{\i t J \eta_{\mu\nu q}}
-  = \bk{\psi\op{\mu}{\nu}\psi} \chi_{\mu\nu}\p{\varphi},
-\end{align}
-and
-\begin{align}
-  \bk{\Psi|s^2\p{t}|\Psi}
-  &= \f1{N^2} \sum_{\mu,\nu} \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2
-  \sum_{p\ne q} e^{\i tJ\p{\eta_{\mu\nu p}-\eta_{\mu\nu q}}} + \f{n}{N} \\
-  &= \sum_{\mu,\nu} \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2
-  \abs{\chi_{\mu\nu}\p{\varphi}}^2 + \f{n-1}{N},
-\end{align}
-where
-\begin{align}
-  \chi_{\mu\nu}\p{\varphi} \equiv \f1N \sum_q e^{\i tJ\eta_{\mu\nu q}}
-  = \f1N \sum_q e^{\i tJ\sin\p{\varphi_{\mu\nu}^-} \sin\p{q+\varphi_{\mu\nu}^+}}.
-\end{align}
-As $N\to\infty$, this sum is well approximated by an integral:
-\begin{align}
-  \chi_{\mu\nu}\p{\varphi} \approx \f1{2\pi} \int_{-\pi}^\pi \dd\theta\,
-  e^{\i tJ\sin\p{\varphi_{\mu\nu}^-} \sin\p{\theta+\varphi_{\mu\nu}^+}}
-  = \J_0\sp{tJ\sin\p{\varphi_{\mu\nu}^-}},
-\end{align}
-where $\J_0$ is the zero-order Bessel function of the first kind.
-To leading order in $N\gg1$, we thus find that
-\begin{align}
-  \bk{\Psi|s_{\mu\nu}\p{t}|\Psi}
-  &\approx \bk{\psi\op{\mu}{\nu}\psi}
-  \J_0\sp{tJ\sin\p{\varphi_{\mu\nu}^-}},
-  \\
-  \bk{\Psi|s^2\p{t}|\Psi}
-  &\approx \sum_{\mu,\nu} \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2
-  \J_0\sp{tJ\sin\p{\varphi_{\mu\nu}^-}}^2.
-  \label{eq:ops_time}
-\end{align}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Limiting cases}
-
-The constant (time independent) terms in \eqref{eq:ops_time} have $\mu=\nu$ (in which case all $\varphi_{\mu\nu}^-=0$), so up to $\O(1/N)$ corrections
-\begin{align}
-  \bk{\Psi|s_{\mu\nu}^{\t{const}}|\Psi}
-  \approx \delta_{\mu\nu} \abs{\bk{\mu|\psi}}^2,
-  &&
-  \bk{\Psi|s^2_{\t{const}}|\Psi}
-  \approx \sum_\mu \abs{\bk{\mu|\psi}}^4 \equiv r_\psi,
-  \label{eq:ops_const}
-\end{align}
-where the quantity $r_\psi$ is sometimes called the inverse participation ratio of $\ket\psi$ with respect to the basis $\set{\ket\mu}$ of states with definite projection onto the $z$ axis.
-Figure \ref{fig:limiting_vals} shows several values of $r_\psi$ for different $n$.
-
-At ``short'' times $t$ for which $tJ\sin\p{I\varphi}\ll1$, we can Taylor expand the Bessel functions in \eqref{eq:ops_time} about $t=0$ to get
-\begin{align}
-  \bk{\Psi|s_{\mu\nu}\p{t}|\Psi}
-  &\approx \bk{\psi\op{\mu}{\nu}\psi}
-  \times \sp{1 - \f14 \p{t J}^2 \sin\p{\varphi_{\mu\nu}^-}^2}, \\
-  \bk{\Psi|s^2\p{t}|\Psi}
-  &\approx 1 - \f12 \p{tJ}^2 \sum_{\mu,\nu}
-  \abs{\bk{\psi\op{\mu}{\nu}\psi}}^2 \sin\p{\varphi_{\mu\nu}^-}^2.
-\end{align}
-In the limit of a small angle $\varphi$ for which $I\varphi\ll1$, we can further simplify
-\begin{align}
-  \bk{\Psi|s_{\mu\nu}\p{t}|\Psi}
-  &\approx \bk{\psi\op{\mu}{\nu}\psi}
-  \times \sp{1 - \f1{16} \p{\f{\mu-\nu}{I}}^2
-    \p{JI\varphi}^2 \times t^2}, \\
-  \bk{\Psi|s^2\p{t}|\Psi}
-  &\approx 1 - \f14 \var_\psi\p{Z} \p{JI\varphi}^2 \times t^2,
-  \label{eq:small_time_angle}
-\end{align}
-where
-\begin{align}
-  \var_\psi\p{Z} \equiv \bk{\psi|Z^2|\psi} - \bk{\psi|Z|\psi}^2,
-  &&
-  Z \equiv \sum_\mu \f{\mu}{I} \op{\mu}.
-\end{align}
-Note that the result in \eqref{eq:small_time_angle} is valid for all $t\ll\p{JI\varphi}^{-1}$, which may be a long time given the small-angle assumption $I\varphi\ll1$.
-Figure \ref{fig:limiting_vals} shows several values of $\var_\psi\p{Z}$ for different $n$.
-
-\begin{figure}
-  \centering
-  \includegraphics{figures/oscillations/limiting_vals.pdf}
-  \caption{Several values of $r_\psi$ and $\var_\psi\p{Z}$ for
-    different $n$.  Computed with respect to the states
-    $\ket\X\equiv\ket{\pi/2,0}$ and $\ket{\DS}$, defined in
-    \eqref{eq:polarized_state} and \eqref{eq:double_state}.}
-  \label{fig:limiting_vals}
-\end{figure}
+Our mean-field theory then simply treats the boson operators $b_{\mu q}$ as complex numbers, with the initial value $b_{\mu q}$ equal to the amplitude of spin $q$ in state $\mu$\footnote{This mean-field theory is mathematically equivalent to a mean-field treatment of spin operators $s_{\mu\nu k}$ with initial conditions set by the corresponding initial-state expectation values.
+The advantage of the Schwinger boson decomposition is that it reduces the number of variables (operators) from $n^2 N$ to $nN$.}.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Strongly-interacting limit of a realistic 1D model}
+\section{Preliminary numerics}
 
-We now consider the strongly-interacting limit of a 1D model, in which the Hamiltonian is
+Here we show some preliminary numerical results from mean-field theory simulations.
+We will examine the dynamical behavior of two quantities: the mean squared magnetization
 \begin{align}
-  H = -\f{U}{N} \v S^\dag\cdot \v S
-  - \f{J}{2} \sum_{\mu,q} \cos\p{q + \mu\varphi} s_{\mu\mu q}
+  s^2 \equiv \f{\bk{\v S}}{N}\c\f{\bk{\v S}}{N}
+  = \sum_{\mu,\nu} \abs{\f{\bk{S_{\mu\nu}}}{N}}^2,
 \end{align}
-with $J\ll U$.
+and the spectrum of the mean reduced single particle density matrix $\bar\rho$ (equivalently, the spectrum of the mean spin matrix $\bk{\v S}/N$).
+We first examine the familiar case
+
+\newpage
+\appendix
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
 \section{Nuclear spin tomography}
+\label{sec:tomography}
 
@@
 To be concrete, we can directly measure projectors $\op{\mu}$ onto states of definite spin projection $\mu$ onto the $z$ axis.
-While the three-laser drive gives us access to a variety of nuclear spin Hamiltonians, most notably it allows us to implement arbitrary SU(2) rotations of the form $e^{-\i\v\theta\c\v S}$, with $\v S\equiv\p{S_\x,S_\y,S_\z}$ a vector of spin operators and $\v\theta\equiv\p{\theta_\x,\theta_\y,\theta_\z}$ an arbitrary rotation vector.
-These rotations essentially allow us to measure projectors $\Pi_{\v v\mu} \equiv \op{\mu_{\v v}}$ onto states of definite spin projection $\mu$ onto an arbitrary quantization axis $\v v\in\SS_2/\ZZ_2$.
+While the three-laser drive gives us access to a variety of nuclear spin Hamiltonians, most notably it allows us to implement SU(2) rotations of the form $e^{-\i\v\theta\c\v s_{\t{3D}}}$, where $\v s_{\t{3D}}\equiv\p{s_\x,s_\y,s_\z}$ is a vector of spin operators that generate rotations in 3D space, and $\v\theta\equiv\p{\theta_\x,\theta_\y,\theta_\z}$ is an arbitrary rotation vector (see Appendix \ref{sec:controls}).
+These rotations essentially allow us to measure projectors $\Pi_{\v v\mu} \equiv \op{\mu_{\v v}}$ onto states of definite spin projection $\mu$ onto an arbitrary quantization axis $\v v\in\SS_2\setminus\ZZ_2$ (where we mod out by $\ZZ_2$ because antipodal points on the sphere correspond to the same set of projectors).
 
@@
 \end{align}
-To summarize: randomly choosing a set $V$ of $2n-1$ axes will induce a set of measurement operators $\Pi_V$ from which we can construct a complete basis $\set{Q^V_\alpha}$ of operators for $\B_n$.
+To summarize: choosing a set $V$ of $2n-1$ random axes will determine a set of measurement operators $\Pi_V$ from which we can construct a complete basis $\set{Q^V_\alpha}$ of operators for $\B_n$.
 However, some choices of axes will be better than others: a good choice of axes $V$ will maximize the amount of information that the corresponding measurement data $M_V=\set{M^V_j}$ will contain about the expectation values $\bk{Q^V_\alpha}_\rho$ used to reconstruct $\rho$.
@@
   A vanishing norm can only occur due to fine-tuned cancellations that make the expectation value of $\sum_j x^V_{\alpha j} \Pi^V_j$ difficult to estimate from noisy data on $\bk{\Pi^V_j}_\rho$.}.
-Rather than choosing axes randomly, we can therefore choose a set of axes that minimizes some cost function $C\p{m_V}$ favoring large eigenvalues $m_V\equiv\set{m^V_\alpha}$.
+Rather than choosing $2n-1$ axes randomly, we can therefore choose a set of axes that minimizes some cost function $C\p{m_V}$ favoring large eigenvalues $m_V\equiv\set{m^V_\alpha}$.
 A reasonable cost function that treats all eigenvalues on equal footing, for example, is
@@
 We leave the details of tailored and adaptive measurement protocols to future work.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{External controls}
+\label{sec:controls}
+
+In addition to letting atoms evolve freely under $H_{\t{spin}}$, we can turn on external control fields to address atoms' internal states.
+Our primary tool here will be three lasers that off-resonantly address an excited electronic state of the atoms.
+Each laser will be indexed by a different integer $m\in\set{1,0,-1}$ that determines the nuclear spin transition $\mu\to\mu+m$ accompanying the electronic $\ket\dn\to\ket\up$ transition upon absorption of a photon.
+Laser $m$ will have a homogeneous drive amplitude $\Omega_m$, and a relative phase $e^{\i\phi_m}$ between adjacent lattice sites.
+All lasers will be detuned by a frequency $\Delta$ away from the electronic $\ket\dn\to\ket\up$ transition.
+Altogether, these lasers induce the drive Hamiltonian
+\begin{align}
+  H_{\t{drive}}^{\t{full}} = \sum_j \sum_{m\in\set{1,0,-1}} \Omega_m
+  \p{e^{\i\phi_m j} s_{mj} \otimes\op{\up}{\dn}_j + \t{h.c.}}
+  + \f{\Delta}{2} \1 \otimes\p{\op\up_j-\op\dn_j},
+  \label{eq:drive_full}
+\end{align}
+where $j$ indexes a lattice site; $s_\pm$ are spin-raising ($+$) and spin-lowering ($-$) operators (containing all appropriate Clebsch-Gordan coefficients); $s_0\coloneqq s_\z$ is a spin-$z$ operator; $\Delta$ is the detuning of the drive lasers from the excited-state transition frequency; and $\1$ is an identity operator (acting on the nuclear spins of all atoms).
+
+In the far-detuned limit $\Delta\gg\Omega_m$, a perturbative treatment of the excited electronic state yields an effective drive Hamiltonian that only addresses nuclear spins.
+Furthermore, if the angles $\phi_m$ satisfy $\phi_m-\phi_{m'}=\p{m-m'}\phi$ for a fixed spin-orbit coupling angle $\phi$ (e.g.~with $\phi_m=m\phi$), then the gauge transformation $c_{j\mu}\to e^{\i\mu\phi j} c_{j\mu}$ makes the drive Hamiltonian spatially homogeneous.
+Altogether, we arrive at a homogeneous effective drive Hamiltonian of the form
+\begin{align}
+  H_{\t{drive}} = \sum_j H_{\t{drive},j}^{\t{single}},
+\end{align}
+where the single-spin Hamiltonian $H_{\t{drive}}^{\t{single}}$ can be conveniently expressed in terms of operators $s_\x,s_\y,s_\z$ that generate rotations in 3D space:
+\begin{align}
+  H_{\t{drive}}^{\t{single}}
+  = h_+ h_- s_\z + h_0 h_+ s_\x
+  + h_0 h_- \p{s_\z s_\x  + s_\x s_\z}
+  - h_0^2\p{s_\z^2-\v s_{\t{3D}}^2} - h_-^2 s_\x^2 - h_+^2 s_\y^2.
+\end{align}
+with $\v s_{\t{3D}}\equiv\p{s_\x,s_\y,s_\z}$.
+Here $h_0\equiv\Omega_0/\sqrt\Delta$ and $h_\pm\equiv\p{\Omega_+\pm\Omega_-}/\sqrt{2\Delta}$ can be tuned independently through the choice of drive amplitudes $\Omega_m$; and $\v s_{\t{3D}}^2=s_\x^2+s_\y^2+s_\z^2=I\p{I+1}$ is a constant.
+Table \ref{tab:drives} shows several drive Hamiltonians that can be implemented with various amplitude-matching conditions.
+We do not expect these controls to be sufficient for implementing arbitrary collective SU($n$) rotations, but identifying the subgroup of SU($n$) generated by these controls could be an interesting problem for future work.
+One notable capability is using pulse sequences to implement SU(2) rotations of the form $e^{-\i\v\theta\c\v s_{\t{3D}}}$ with an arbitrary rotation vector $\v\theta\equiv\p{\theta_\x,\theta_\y,\theta_\z}$.
+In addition to enabling us to perform arbitrary 3D rotations, these collective rotations enable us to measure all components of the mean collective spin matrix $\bk{\v S}$, as discussed in Appendix \ref{sec:tomography}.
+
+\begin{table}
+  \centering
+  \caption{Drive Hamiltonians that can be implemented with various amplitude-matching conditions.}
+  \begin{tabular}{c|c|c}
+    Condition & $\p{h_0,h_+,h_-}$
+    & $H_{\t{drive}}^{\t{single}}$
+    \\ \hline\hline
+    $\Omega_+=\Omega_-=0$ & $\p{1,0,0}$ & $-s_\z^2 + \v s_{\t{3D}}^2$
+    \\ \hline
+    $\Omega_0=0$,  $\Omega_+=+\Omega_-$ & $\p{0,1,0}$ & $-s_\y^2$
+    \\ \hline
+    $\Omega_0=0$,  $\Omega_+=-\Omega_-$ & $\p{0,0,1}$ & $-s_\x^2$
+    \\ \hline
+    $\Omega_0=\Omega_\mp=0$ & $\p{0,1,\pm1}$
+    & $\pm s_\z + s_\z^2 - \v s_{\t{3D}}^2$
+    \\ \hline
+    $\sqrt{2}\Omega_0=\pm\Omega_+=\pm\Omega_-$
+    & $\p{1,\pm1,0}$ & $\pm s_\x + s_\x^2$
+    \\ \hline
+    $\sqrt{2}\Omega_0=\pm\Omega_+=\mp\Omega_-$
+    & $\p{1,0,\pm1}$ & $\pm\p{s_\z s_\x+s_\x s_\z} + s_\y^2$
+    \\ \hline
+    $\Omega_0=\pm\Omega_+=\pm\Omega_-$ & $\p{1,\pm\sqrt{2},0}$
+    & $\pm\sqrt{2} s_\x + \p{s_\x^2 - s_\y^2}$
+    \\ \hline
+    $\Omega_0=\pm\Omega_+=\mp\Omega_-$ & $\p{1,0,\pm\sqrt{2}}$
+    & $\pm\sqrt{2}\p{s_\z s_\x + s_\x s_\z} - \p{s_\x^2-s_\y^2}$
+  \end{tabular}
+  \label{tab:drives}
+\end{table}
+
+As a final comment, we note that we can, in principle, address nuclear spins directly using external magnetic fields.
+We can also make use of a differential magnetic (Zeeman) splitting, whereby the strength of nuclear spin coupling to an external magnetic field depends on the electronic state of the atoms.
+These controls suffer from the fact that nuclear spin coupling to external magnetic fields is typically weak.
+Even so, these controls are worth exploring in the future.
 

### 2020-09-17 sun_phases/sun_dynamics_notes.tex
final changes before sending notes out to collaborators added appendix deriving Schwinger boson equations of motion

@@
 \newcommand{\Z}{\text{Z}}
-\newcommand{\DCS}{\text{DCS}}
+\newcommand{\BTBX}{\text{BTB-X}}
 
 \newcommand{\B}{\mathcal{B}}
+\renewcommand{\H}{\mathcal{H}}
+\renewcommand{\P}{\mathcal{P}}
 
@@
 \newcommand{\ZZ}{\mathbb{Z}}
+
+\DeclareMathOperator{\sign}{sign}
+
+%%% figures
+\usepackage{graphicx} % for figures
+\graphicspath{{./figures/}} % set path for all figures
+
+% to place figures in the correct section
+\usepackage[section]{placeins}
 
@@
 
-\title{Collective SU($n$) spin dynamics}%
+\title{SU($n$) spin model dynamics}%
 \author{Michael A. Perlin}%
@@
 
-Here, we will consider a particular collective spin model resembling that used to study BCS superconductivity.
-We begin by briefly deriving the spin model, before discussing some of its symmetries, examples of initial states that can be prepared, and additional experimental controls.
+Here, we will consider a particular multilevel spin model resembling that used to study BCS superconductivity.
+We begin by briefly deriving the spin model, before discussing some of its symmetries, examples of initial states that can be prepared experimentally, and preliminary numerical mean-field numerics.
+Additional information about some experimental capabilities and controls are provided in appendices.
 
@@
 \begin{align}
-  H_{\t{bare}} = -\f{J}{2} \sum_{j,\mu}
+  H_{\t{bare}} = -J \sum_{j,\mu}
   \p{c_{j\mu}^\dag c_{j+1,\mu} + \t{h.c.}}
-  + U \sum_{j,\mu,\nu} c_{j\mu}^\dag c_{j\mu} c_{j\nu}^\dag c_{j\nu},
+  + \f{U}{2} \sum_{j,\mu,\nu} c_{j\mu}^\dag c_{j\nu}^\dag c_{j\nu} c_{j\mu},
   \label{eq:bare_sites}
 \end{align}
-where $c_{j\mu}$ annihilates a fermionic atom on lattice site $j$ with nuclear spin projection $\mu\in\set{I,I-1,\cdots,-I}$ onto a quantization axis; $I\equiv\p{n-1}/2$ is the total nuclear spin of each atom (e.g.~$I=9/2$ in the case of Sr-87 with $n=10$ nuclear spin levels); $J$ is a tunneling rate; and $U$ is an on-site interaction strength.
+where $c_{j\mu}$ annihilates a fermionic atom on lattice site $j$ with nuclear spin projection $\mu\in\set{I,I-1,\cdots,-I}$ onto a quantization axis; $I\equiv\frac{n-1}{2}$ is the total nuclear spin of each atom (e.g.~$I=9/2$ in the case of Sr-87 with $n=10$ nuclear spin levels); $J$ is the nearest-neighbor tunneling rate; and $U$ is the two-body on-site interaction strength.
 In terms of quasi-momentum modes, the same Hamiltonian reads
 \begin{align}
-  H_{\t{bare}} = -J \sum_{q,\mu} \cos q\, c_{q\mu}^\dag c_{q\mu}
-  + \f{U}{L} \sum_{\substack{k,\ell,p,q\\\mu,\nu}}
-  \delta_{k+\ell,p+q}\, c_{k\mu}^\dag c_{p\mu} c_{\ell\nu}^\dag c_{q\nu},
+  H_{\t{bare}} = -2J \sum_{q,\mu} \cos q\, c_{q\mu}^\dag c_{q\mu}
+  + \f{U}{2L} \sum_{\substack{k,\ell,p,q\\\mu,\nu}}
+  \delta_{k+\ell,p+q}\, c_{k\mu}^\dag c_{\ell\nu}^\dag c_{p\nu} c_{q\mu},
   \label{eq:bare_momenta}
@@
 where $L$ is the number of lattice sites; $k,\ell,p,q$ index quasi-momenta; and $\delta_{k+\ell,p+q}$ enforces momentum conservation.
-If interactions are sufficiently weak, $U\lesssim J$, then we can neglect mode-changing collisions and only keep interaction terms in which the final momenta are the same as the initial momenta, $\set{p,q}=\set{k,\ell}$.
-In this case, the ``direct'' terms with $\p{p,q}=\p{k,\ell}$ turn out to be trivial (proportional to the total atom number squared, $N^2$), leaving only ``exchange'' terms with $\p{p,q}=\p{\ell,k}$.
+If interactions are sufficiently weak, $U\lesssim J$, then we can neglect mode-changing collisions and only keep interaction terms in which the final momenta are the same as the initial momenta, $\set{k,\ell}=\set{p,q}$.
+In this case, terms with $\p{k,\ell}=\p{q,p}$ turn out to be trivial (proportional to the total atom number squared, $N^2$), leaving only terms with $\p{k,\ell}=\p{p,q}$.
 In total, the operator content of the interactions can be written in the form
@@
   \sum_{\substack{k,\ell,p,q\\\mu,\nu}}
-  \delta_{k+\ell,p+q}\, c_{k\mu}^\dag c_{p\mu} c_{\ell\nu}^\dag c_{q\nu}
-  \to \sum_{k,\ell,\mu,\nu} c_{k\mu}^\dag c_{\ell\mu}
-  c_{\ell\nu}^\dag c_{k\nu}
+  \delta_{k+\ell,p+q}\, c_{k\mu}^\dag c_{\ell\nu}^\dag c_{p\nu} c_{q\mu}
+  \to \sum_{k,\ell,\mu,\nu} c_{k\mu}^\dag c_{\ell\nu}^\dag
+  c_{k\nu} c_{\ell\mu}
   = -\sum_{k,\ell,\mu,\nu} s_{\mu\nu k} s_{\nu\mu\ell}
@@
 \end{align}
-where we define spin operators $s_{\mu\nu k}\equiv c_{\mu k}^\dag c_{\nu k}$; $\v s_k=\sum_{\mu,\nu}s_{\mu\nu k}\op{\mu}{\nu}$ is an operator-valued spin matrix for mode $k$; $\v S\equiv\sum_k \v s_k$ is a collective spin matrix; and $\v A\c\v B\equiv\tr\p{\v A^\dag \v B}$ is an inner product of spin matrices (whose result is, again, operator-valued).
+where we define spin operators $s_{\mu\nu k}\equiv c_{k\mu}^\dag c_{k\nu}$; $\v s_k=\sum_{\mu,\nu}s_{\mu\nu k}\op{\mu}{\nu}$ is an operator-valued spin matrix for mode $k$; $\v S\equiv\sum_k \v s_k$ is a collective spin matrix; and $\v A\c\v B\equiv\tr\p{\v A^\dag \v B}$ is an inner product of spin matrices (whose result is, again, operator-valued).
 Putting everything together, we arrive at a spin Hamiltonian of the form
@@
   H_{\t{spin}}^{\t{bare}}
-  = -\f{U}{L} \v S \c\v S - J \sum_{q,\mu} \cos q\, s_{\mu\mu q}.
+  = -\f{U}{2L} \v S \c\v S - 2J \sum_{q,\mu} \cos q\, s_{\mu\mu q}.
   \label{eq:bare_spin}
@@
 \begin{align}
-  H_{\t{spin}} = -\f{U}{L} \v S \c\v S
-  - J \sum_{q,\mu} \cos\p{q+\mu\phi}\, s_{\mu\mu q}.
+  H_{\t{spin}} = -\f{U}{2L} \v S \c\v S
+  - 2J \sum_{q,\mu} \cos\p{q+\mu\phi}\, s_{\mu\mu q}.
   \label{eq:spin_cos}
 \end{align}
-The resulting dispersion relation is shown in Figure \ref{fig:dispersion}.
+The resulting dispersion relation is shown in Figure \ref{fig:disps}.
 This Hamiltonian can be written in the simplified form
 \begin{align}
-  H_{\t{spin}} = -\f{U}{L} \v S\c\v S
-  - J \sum_q \p{\cos q\, w_{\phi,q}^+ - \sin q\, w_{\phi,q}^-},
+  H_{\t{spin}} = -\f{U}{2L} \v S\c\v S
+  - 2J \sum_q \p{\cos q\, w_{\phi,q}^+ - \sin q\, w_{\phi,q}^-},
   \label{eq:spin_pair}
@@
   \centering
-  \includegraphics{figures/sun_dispersion.pdf}
-  \caption{Single-particle dispersion relations $E_{q\mu}=-J\cos\p{q+\mu\phi}$ for the case of $n=6$ (nuclear spin $I=5/2$) and different spin-orbit coupling angles $\phi$.
+  \includegraphics{sun_dispersion.pdf}
+  \caption{Single-particle dispersion relations $E_{q\mu}=-2J\cos\p{q+\mu\phi}$ for the case of $n=6$ (nuclear spin $I=5/2$) and different spin-orbit coupling angles $\phi$.
     Legend indicates the spin projection $\mu$ onto the $z$.
-    Note that these dispersion relations are ``maximally spread'' when $\phi=2\pi/n$ (as with $\phi=\pi/3$ above), and that degeneracies occur at all $q$ when $\phi=2\pi/m$ for any $m\in\set{1,2,\cdots,n-1}$.}
-  \label{fig:dispersion}
+    Note that these dispersion relations are ``maximally spread'' when $\phi=2\pi/n$ (as with $\phi=\pi/3$ above), and that some of these dispersions overlap (implying degeneracies at all $q$) when $\phi=2\pi/m$ for any $m\in\set{1,2,\cdots,n-1}$.}
+  \label{fig:disps}
 \end{figure}
@@
 \begin{align}
-  H_{\t{spin}}^{(n=2)} = -\f{U}{L} \v S\c\v S
-  + J \sin\p{\phi/2} \sum_q \sin q\, \sigma_{\z,q}
-  - J \cos\p{\phi/2} \sum_q \cos q\, \1_q,
+  H_{\t{spin}}^{(n=2)} = -\f{U}{2L} \v S\c\v S
+  + 2J \sin\p{\phi/2} \sum_q \sin q\, \sigma_{\z,q}
+  - 2J \cos\p{\phi/2} \sum_q \cos q\, \1_q,
 \end{align}
@@
 \begin{align}
-  H_{\t{spin}} = -\f{U}{L} \v S\c\v S
-  + J \phi \sum_q \sin q\, s_{\z,q}
-  - J \sum_q \cos q\, \1_q + O(\phi^2),
+  H_{\t{spin}} = -\f{U}{2L} \v S\c\v S
+  + 2J \phi \sum_q \sin q\, s_{\z,q}
+  - 2J \sum_q \cos q\, \1_q + O(\phi^2),
   &&
@@
 \begin{align}
-  H_{\t{spin}}^{(\phi=\pi)} = -\f{U}{L} \v S\c\v S
-  + J \sum_q \sin q\, s_{\a,q},
+  H_{\t{spin}}^{(\phi=\pi)} = -\f{U}{2L} \v S\c\v S
+  + 2J \sum_q \sin q\, s_{\a,q},
   &&
@@
 Needless to say, the single-body terms in \eqref{eq:spin_pair} explicitly break the SU($n$) symmetry of the interactions.
-Even so, the fact that all single-body terms are diagonal with respect to nuclear spin projection $\mu$ implies that the overall Hamiltonian conserves all collective spin operators of the form $S_{\mu\mu}\equiv\sum_q s_{\mu\mu q}$, as well as all products thereof, which is equivalent to conserving all powers of $S_\z$ in the case of SU(2).
-The enormous number of these conserved quantities\footnote{Altogether, there are essentially ${N+n-1 \choose n-1}\sim N^{n-1}$ such quantities.} suggests that the spin Hamiltonian in \eqref{eq:spin_pair} is integrable, although the precise form of an appropriate Bethe ansatz or Lax formulation is unclear.
-The situation is somewhat simpler when $\phi\ll1$ or $\phi=\pi$, in which case the single-body terms are built from a single spin operator (namely, $s_\z$ or $s_\a$).
+Even so, the fact that all single-body terms are diagonal with respect to nuclear spin projection $\mu$ implies that the overall Hamiltonian conserves all collective spin operators of the form $S_{\mu\mu}\equiv\sum_q s_{\mu\mu q}$, as well as all products thereof, which is equivalent to conserving all powers of $S_\z$ in the case of SU(2)\footnote{This list of conserved quantities is not exhaustive, and in particular a specific choice of initial state may result in additional conserved quantities.}.
+The fact that the Hamiltonian is diagonal in the ``computational'' basis of all spins (i.e.~the basis of their spin projection onto the $z$ axis) and suggests that the spin Hamiltonian in \eqref{eq:spin_pair} might be integrable, although the precise form of an appropriate Bethe ansatz or Lax formulation is unclear.
+When $\phi\ll1$ or $\phi=\pi$, however, the single-body terms of the spin Hamiltonian are built from a single spin operator (namely, $s_\z$ or $s_\a$), so extending the Bethe ansatz or Lax formulation of the SU(2) case to SU($n$) is relatively straightforward.
 
-It is worth mentioning at this point that we can, in principle, perform full tomography on the average single-body reduced density operator $\bar\rho$, which is equivalent to measuring all components of the mean collective spin matrix $\bk{\v S}\equiv\sum_{\mu,\nu}\bk{S_{\mu\nu}}\op{\mu}{\nu}$ (see Appendix \ref{sec:tomography} for additional details).
-The squared ``magnetization'' $\bk{\v S}\c\bk{\v S}=\sum_{\mu,\nu}\abs{\bk{S_{\mu\nu}}}^2$, which is essentially equal to the purity of $\bar\rho$, may therefore be an interesting ``unbiased'' dynamical order parameter to consider.
+It is worth mentioning at this point that we can, in principle, perform full tomography on the average single-body reduced density operator $\bar\rho$, which is equivalent to measuring all components of the mean collective spin matrix $\bk{\v S}\equiv\sum_{\mu,\nu}\bk{S_{\mu\nu}}\op{\mu}{\nu}$ (see Appendix \ref{sec:tomography} for additional details)\footnote{Fluctuations in $\v S$, and thereby $\bk{\v S\c\v S}$, are also experimentally accessible in principle, but this quantity is not accessible via mean-field theory.}.
+The squared mean ``magnetization'' $\bk{\v S}\c\bk{\v S}=\sum_{\mu,\nu}\abs{\bk{S_{\mu\nu}}}^2$, which is essentially equal to the purity of $\bar\rho$, is therefore an experimentally accessible  ``unbiased'' dynamical order parameter that we can consider.
 In addition, diagonalizing $\bar\rho$ (equivalently, $\bk{\v S}$), provides a ``natural'' basis in which to examine collective observables.
@@
 Using external control fields (discussed in the Section \ref{sec:controls}), we can rotate the state $\ket\Z$ to orient it along any axis in 3D space.
-External control fields also allow us to prepare states analogous to $\p{\ket\up+\ket\dn}^{\otimes N}$ in that each spin is in a coherent superposition of states polarized along different directions, e.g.~$\p{\ket{I}+\ket{-I}}^{\otimes N}$.
-The possibilities to prepare $N$-fold product states of the form $\ket\psi^{\otimes N}$ are enormous.
+External control fields also allow us to prepare product states in which each spin is in a coherent superposition of states polarized along different directions, e.g.~$\p{\ket{+I}+\ket{-I}}^{\otimes N}$.
+Generally speaking, the possibilities to prepare $N$-fold product states of the form $\ket\psi^{\otimes N}$ are enormous.
 
@@
 
-The dynamics induced by the spin Hamiltonian in \eqref{eq:spin_cos} and \eqref{eq:spin_pair} are efficiently simulable when $U=0$ or $U\gg JI\phi$.
+Given an initial product state, the dynamics induced by the spin Hamiltonian in \eqref{eq:spin_cos} and \eqref{eq:spin_pair} are (in principle) classically simulable when $U=0$ or $U\gg JI\phi$.
 In order to explore the entirety of parameter space, however, we turn to mean-field theory.
-To this end, we first represent individual spin operators by Schwinger bosons, $s_{\mu\nu q}\coloneqq b_{\mu q}^\dag b_{\nu q}$.
+To this end, we first decompose individual spin operators into Schwinger bosons, $s_{\mu\nu q}\coloneqq b_{\mu q}^\dag b_{\nu q}$.
 Up to an overall constant, our spin model can then be written in the form
 \begin{align}
-  H_{\t{boson}} = - \f{U}{L} \sum_{\substack{\mu,\nu\\p\ne q}}
+  H_{\t{boson}} = - \f{U}{L} \sum_{\substack{\mu,\nu\\p<q}}
   b_{\mu p}^\dag b_{\nu p} b_{\nu q}^\dag b_{\mu q}
-  - J \sum_{\mu,q} \cos\p{q+\mu\phi} b_{\mu q}^\dag b_{\mu q}.
+  - 2J \sum_{\mu,q} \cos\p{q+\mu\phi} b_{\mu q}^\dag b_{\mu q}.
 \end{align}
-With a bit of algebra, we can work out that the Heisenberg equations of motion for individual boson operators are
+With a bit of algebra (see Appendix \ref{sec:bosons}), we can work out that the Heisenberg equations of motion for the individual boson operators are
 \begin{align}
   \i \d_t b_{\mu q} = \sp{b_{\mu q},H}
-  = -2\, \f{U}{L} \sum_{\nu,p} b_{\nu p}^\dag b_{\mu p} b_{\nu q}
-  - J \cos\p{q+\mu\phi} b_{\mu q}.
+  = -\f{U}{L} \sum_{\nu,p} b_{\nu p}^\dag b_{\mu p} b_{\nu q}
+  - 2J \cos\p{q+\mu\phi} b_{\mu q}.
 \end{align}
@@
 \section{Preliminary numerics}
+\label{sec:numerics}
 
-Here we show some preliminary numerical results from mean-field theory simulations.
-We will examine the dynamical behavior of two quantities: the mean squared magnetization
+Here we show some preliminary numerical results from mean-field theory simulations with $N=20$ spins.
+We apologize in advance for the crude presentation of these results.
+The purpose here is mostly to give sense of what sorts of quantities we looked at so far, rather than to provide a detailed and organized summary of our findings (we are still exploring possible initial states, observables, and dynamical behaviors, so such a summary does not yet exist).
+
+We will generally assume unit filling, setting $L=N$, and here consider spin-orbit coupling angles $\phi\in\set{\pi/2,\pi}$\footnote{In retrospect, we note that these choices of spin-orbit coupling angles are not ideal for examining ``generic'' behavior, because they yield degenerate dispersions, as discussed in the caption of Figure \ref{fig:disps}.}.
+We will also consider two types of initial states: a fully polarized state $\ket\X$ in which all spins point along the $x$ axis, and a ``back-to-back'' state $\ket\BTBX$ in which half of the spins are polarized along $+x$, and half along $-x$.
+Specifically, indexing spins by their associated quasi-momentum mode $q\in[-\pi,\pi)$, the initial state $\ket\BTBX$ has spins with $q\ge0$ ($q<0$) initially polarized along $+x$ ($-x$).
+In this way, the initial state $\ket\BTBX$ is correlated with $\sim\sin q$ terms of the spin Hamiltonian in \eqref{eq:spin_pair}.
+
+For each initial state, $\ket\X$ or $\ket\BTBX$, and each spin-orbit coupling angle $\phi\in\set{\pi/2,\pi}$, we will examine how the dynamical behavior of a few quantities depends on the strength of single-particle terms relative to interactions, $J/U$.
+In particular, we will examine the mean squared magnetization
 \begin{align}
@@
 \end{align}
-and the spectrum of the mean reduced single particle density matrix $\bar\rho$ (equivalently, the spectrum of the mean spin matrix $\bk{\v S}/N$).
-We first examine the familiar case
+and the spectrum of the mean reduced single particle density matrix $\bar\rho$ (equivalently, the spectrum of the mean spin matrix $\bk{\v S}/N=\sum_{\mu,\nu}\bk{S_{\mu\nu}}/N \op{\mu}{\nu}$).
+
+Figures \ref{fig:ss_spect_n2_X_p0.5}--\ref{fig:ss_spect_n2_XX_p1.0} show the time-series of $s^2(t)$, its power spectral density $\widetilde{s^2}(\omega)$, and time-series spectrum of $\bar\rho(t)$ for the case of $n=2$, and with values of $J/U$ between $10^{-2}$ and $10^1$.
+In short, these figures all tell the same story: there is a ``collective'' phase when $J/U\lesssim1$, and of a ``single-particle'' phase when $J/U\gtrsim1$.
+No surprises here.
+
+Figure \ref{fig:ss_spect_n6_X_p0.5}--\ref{fig:ss_spect_n6_XX_p1.0} show the same plots for the case of $n=6$.
+Unsurprisingly, here again there appear to be (at least) two phases, at $J/U\lesssim1$ vs.~$J/U\gtrsim1$.
+An interesting observation, however, is that (for both initial states we consider) the spectrum of $\bar\rho$ appears to have only four non-zero eigenvalues when $\phi=\pi/2$, and two non-zero eigenvalues when $\phi=\pi$.
+The case of $\phi=\pi$ happens to be special, as it turns out that (for the initial states we consider) it can be mapped exactly onto a spin model with $n=2$, regardless of the actual value of $n$.
+The general case of $\phi\ne\pi$, however, seems to yield four eigenvalues independent of the actual value of $\phi$.
+One might suspect then that $\bar\rho$ is effectively a density operator on some fixed (time-independent) four-dimensional subspace of $\H_n$ (i.e.~the Hilbert space of an $n$-level quantum system), but that does not appear to be the case: decomposing $\bar\rho\equiv\sum_\lambda\lambda\op\lambda$, we find that the projector $\P_{\bar\rho}\equiv\sum_{\lambda>0}\op\lambda$ is time dependent\footnote{To clarify: $\P_{\bar\rho}$ is time dependent even if you only consider times when $\tr\P_{\bar\rho}=4$.
+  It should be possible to understand this fact in the non-interacting limit $U/J\to0$.}!
+This finding, as well as others (such as the possible existence of a third phase in between those at $J/U\ll1$ and $J/U\gg1$) have yet to be understood and explored in full.
+
+%%% n = 2, X
+
+\begin{figure}
+  \centering \includegraphics{oscillations/time_series/ss_spect_n2_N20_X_p0.50.pdf}
+  \caption{Time-series and power spectral density of $s^2$ (left and middle columns), as well as the spectrum of $\bar\rho$ (right column), for $N=20$ spins with $n=2$ initially in $\ket\X$, with spin-orbit coupling angle $\phi=\pi/2$.}
+  \label{fig:ss_spect_n2_X_p0.5}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics{oscillations/time_series/ss_spect_n2_N20_X_p1.00.pdf}
+  \caption{Time-series and power spectral density of $s^2$ (left and middle columns), as well as the spectrum of $\bar\rho$ (right column), for $N=20$ spins with $n=2$ initially in $\ket\X$, with spin-orbit coupling angle $\phi=\pi$.}
+  \label{fig:ss_spect_n2_X_p1.0}
+\end{figure}
+
+%%% n = 2, XX
+
+\begin{figure}
+  \centering \includegraphics{oscillations/time_series/ss_spect_n2_N20_XX_p0.50.pdf}
+  \caption{Time-series and power spectral density of $s^2$ (left and middle columns), as well as the spectrum of $\bar\rho$ (right column), for $N=20$ spins with $n=2$ initially in $\ket\BTBX$, with spin-orbit coupling angle $\phi=\pi/2$.}
+  \label{fig:ss_spect_n2_XX_p0.5}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics{oscillations/time_series/ss_spect_n2_N20_XX_p1.00.pdf}
+  \caption{Time-series and power spectral density of $s^2$ (left and middle columns), as well as the spectrum of $\bar\rho$ (right column), for $N=20$ spins with $n=2$ initially in $\ket\BTBX$, with spin-orbit coupling angle $\phi=\pi$.}
+  \label{fig:ss_spect_n2_XX_p1.0}
+\end{figure}
+
+%%% n = 6, X
+
+\begin{figure}
+  \centering \includegraphics{oscillations/time_series/ss_spect_n6_N20_X_p0.50.pdf}
+  \caption{Time-series and power spectral density of $s^2$ (left and middle columns), as well as the spectrum of $\bar\rho$ (right column), for $N=20$ spins with $n=6$ initially in $\ket\X$, with spin-orbit coupling angle $\phi=\pi/2$.}
+  \label{fig:ss_spect_n6_X_p0.5}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics{oscillations/time_series/ss_spect_n6_N20_X_p1.00.pdf}
+  \caption{Time-series and power spectral density of $s^2$ (left and middle columns), as well as the spectrum of $\bar\rho$ (right column), for $N=20$ spins with $n=6$ initially in $\ket\X$, with spin-orbit coupling angle $\phi=\pi$.}
+  \label{fig:ss_spect_n6_X_p1.0}
+\end{figure}
+
+%%% n = 6, XX
+
+\begin{figure}
+  \centering \includegraphics{oscillations/time_series/ss_spect_n6_N20_XX_p0.50.pdf}
+  \caption{Time-series and power spectral density of $s^2$ (left and middle columns), as well as the spectrum of $\bar\rho$ (right column), for $N=20$ spins with $n=6$ initially in $\ket\BTBX$, with spin-orbit coupling angle $\phi=\pi/2$.}
+  \label{fig:ss_spect_n6_XX_p0.5}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics{oscillations/time_series/ss_spect_n6_N20_XX_p1.00.pdf}
+  \caption{Time-series and power spectral density of $s^2$ (left and middle columns), as well as the spectrum of $\bar\rho$ (right column), for $N=20$ spins with $n=6$ initially in $\ket\BTBX$, with spin-orbit coupling angle $\phi=\pi$.}
+  \label{fig:ss_spect_n6_XX_p1.0}
+\end{figure}
 
@@
 To be concrete, we can directly measure projectors $\op{\mu}$ onto states of definite spin projection $\mu$ onto the $z$ axis.
-While the three-laser drive gives us access to a variety of nuclear spin Hamiltonians, most notably it allows us to implement SU(2) rotations of the form $e^{-\i\v\theta\c\v s_{\t{3D}}}$, where $\v s_{\t{3D}}\equiv\p{s_\x,s_\y,s_\z}$ is a vector of spin operators that generate rotations in 3D space, and $\v\theta\equiv\p{\theta_\x,\theta_\y,\theta_\z}$ is an arbitrary rotation vector (see Appendix \ref{sec:controls}).
+While the three-laser drive gives us access to a variety of nuclear spin Hamiltonians, most notably it allows us to implement SU(2) rotations of the form $e^{-\i\v\theta\c\v s_{\t{3D}}}$, where $\v s_{\t{3D}}\equiv\p{s_\x,s_\y,s_\z}$ is a vector of spin operators that generate rotations of an $n$-level spin (with total spin $I=\frac{n-1}{2}$) in 3D space, and $\v\theta\equiv\p{\theta_\x,\theta_\y,\theta_\z}$ is an arbitrary rotation vector (see Appendix \ref{sec:controls}).
 These rotations essentially allow us to measure projectors $\Pi_{\v v\mu} \equiv \op{\mu_{\v v}}$ onto states of definite spin projection $\mu$ onto an arbitrary quantization axis $\v v\in\SS_2\setminus\ZZ_2$ (where we mod out by $\ZZ_2$ because antipodal points on the sphere correspond to the same set of projectors).
@@
 \end{align}
-As it turns out, numerical experiments with $n\le20$ suggest that in order to perfurm full tomography of $\rho$, it suffices to measure all projectors $\Pi_{\v v\mu}$ onto states of definite spin $\mu$ along $2n-1$ different axes $\v v$.
+As it turns out, numerical experiments with $n\le20$ suggest that in order to perfurm full tomography of $\rho$, it suffices to measure all projectors $\Pi_{\v v\mu}$ onto states of definite spin $\mu$ along $2n-1$ different axes $\v v$\footnote{It would be interesting to develop an actual (analytical) understanding of this numerical discovery.}.
 The question remains: which axes should we choose?
@@
 where the sum is implicitly performed over the $n^2$ largest eigenvalues of $B_V$ (as all other eigenvalues are guaranteed to be zero).
-Note that any information about $\rho$, obtained either from prior knowledge or preliminary measurement data, can be used to construct an even better choice of measurement axes, leading to the possibility of adaptive measurement protocols \cite{pereira2018adaptive} that are more efficient in terms of the number of measurements required to estimate $\rho$ to a fixed precision.
+Note that any information about $\rho$, obtained either from prior knowledge or preliminary measurement data, can be used to construct an even better choice of measurement axes, leading to the possibility of tailored or adaptive measurement protocols \cite{pereira2018adaptive} that are more efficient in terms of the number of measurements required to estimate $\rho$ to a fixed precision.
 We leave the details of tailored and adaptive measurement protocols to future work.
@@
   \p{e^{\i\phi_m j} s_{mj} \otimes\op{\up}{\dn}_j + \t{h.c.}}
-  + \f{\Delta}{2} \1 \otimes\p{\op\up_j-\op\dn_j},
+  + \f{\Delta}{2}\, \1 \otimes\p{\op\up_j-\op\dn_j},
   \label{eq:drive_full}
@@
 \end{align}
-where the single-spin Hamiltonian $H_{\t{drive}}^{\t{single}}$ can be conveniently expressed in terms of operators $s_\x,s_\y,s_\z$ that generate rotations in 3D space:
+where the single-spin Hamiltonian $H_{\t{drive}}^{\t{single}}$ can be conveniently expressed in terms of operators $s_\x,s_\y,s_\z$ that generate rotations of an $n$-level spin (with total spin $I=\frac{n-1}{2}$) in 3D space:
 \begin{align}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Schwinger boson equations of motion}
+\label{sec:bosons}
+
+Here we consider a quadratic spin Hamiltonian of the general form
+\begin{align}
+  H = \sum_{\substack{\mu,\nu,\rho,\sigma\\j<k}}
+  h^{\mu\nu j}_{\rho\sigma k} s_{\mu\nu j} s_{\rho\sigma k}
+  + \sum_{\mu,\nu,j} \epsilon_{\mu\nu j} s_{\mu\nu j},
+  \label{eq:spin}
+\end{align}
+where $\mu,\nu$ index orthogonal states of an $n$-level spin; $j,k\in\ZZ_N$ index one of $N$ spins; $h^{\mu\nu j}_{\rho\sigma k}$ and $\epsilon_{\mu\nu j}$ are scalars; and $s_{\mu\nu j}\coloneqq\op{\mu}{\nu}_j$ is a transition operator for spin $j$.
+We can write this Hamiltonian using Schwinger bosons as
+\begin{align}
+  H = \sum_{\substack{\mu,\nu,\rho,\sigma\\j<k}}
+  h^{\mu\nu j}_{\rho\sigma k}
+  b_{\mu j}^\dag b_{\nu j} b_{\rho k}^\dag b_{\sigma k}
+  + \sum_{\mu,\nu,j} \epsilon_{\mu\nu j} b_{\mu j}^\dag b_{\nu j},
+\end{align}
+where $b_{\mu j}$ annihilates a boson of type $\mu$ on site $j$.
+The Heisenberg equations of motion for these operators are\footnote{The Hamiltonian in \eqref{eq:spin} only defines $h^{\mu\nu j}_{\rho\sigma k}$ for $j<k$.
+  To simplify expressions in the remainder of this work, we therefore additionally define $h^{\rho\sigma k}_{\mu\nu j}=h^{\mu\nu j}_{\rho\sigma k}$ and $h^{\mu\nu j}_{\rho\sigma j}=0$.}
+\begin{align}
+  \i \d_t b_{\alpha\ell} = \sp{b_{\alpha\ell}, H}
+  &= \sum_{\substack{\mu,\nu,\rho,\sigma\\j<k}}
+  h^{\mu\nu j}_{\rho\sigma k}
+  \sp{b_{\alpha\ell}, b_{\mu j}^\dag b_{\nu j} b_{\rho k}^\dag b_{\sigma k}}
+  + \sum_{\mu,\nu,j} \epsilon_{\mu\nu j}
+  \sp{b_{\alpha\ell}, b_{\mu j}^\dag b_{\nu j}} \\
+  &= \sum_{\mu,\nu,\rho,\sigma,k} h^{\mu\nu\ell}_{\rho\sigma k}
+  \sp{b_{\alpha\ell}, b_{\mu\ell}^\dag b_{\nu\ell}}
+  b_{\rho k}^\dag b_{\sigma k}
+  + \sum_{\mu,\nu} \epsilon_{\mu\nu\ell}
+  \sp{b_{\alpha\ell}, b_{\mu\ell}^\dag b_{\nu\ell}} \\
+  &= \sum_{\mu,\nu} \p{\sum_{\rho,\sigma,k}
+    h^{\mu\nu\ell}_{\rho\sigma k} b_{\rho k}^\dag b_{\sigma k}
+    + \epsilon_{\mu\nu\ell}}
+  \sp{b_{\alpha\ell}, b_{\mu\ell}^\dag b_{\nu\ell}}
+\end{align}
+where
+\begin{align}
+  \sp{b_{\alpha\ell}, b_{\mu\ell}^\dag b_{\nu\ell}}
+  = \delta_{\mu\alpha} \delta_{\nu\alpha} b_{\alpha\ell}
+  + \delta_{\mu\alpha} \p{1-\delta_{\nu\alpha}} b_{\nu\ell}
+  = \delta_{\mu\alpha} b_{\nu\ell},
+\end{align}
+so
+\begin{align}
+  \i \d_t b_{\alpha\ell}
+  = \sum_\nu \p{\sum_{\rho,\sigma,k}
+    h^{\alpha\nu\ell}_{\rho\sigma k} b_{\rho k}^\dag b_{\sigma k}
+    + \epsilon_{\alpha\nu\ell}} b_{\nu\ell}.
+\end{align}
+
 \bibliography{sun_dynamics_notes.bib}

### 2020-09-17 sun_phases/notes.tex
removed content from notes.tex that was moved into sun_dynamics_notes.tex

@@
 \maketitle
-
-We first consider a general quadratic spin Hamiltonian of the form
-\begin{align}
-  H = \f1N \sum_{\substack{\mu,\nu,\rho,\sigma\\i<j}}
-  h^{\mu\nu i}_{\rho\sigma j} s_{\mu\nu i} s_{\rho\sigma j}
-  + \sum_{\mu,\nu,i} \epsilon_{\mu\nu i} s_{\mu\nu i},
-  \label{eq:spin}
-\end{align}
-where $\mu,\nu\in\set{I,I-1,\cdots,-I}$ with $I\equiv\p{n-1}/2$ index orthogonal states of an $n$-level spin; $i,j\in\ZZ_N$ index one of $N$ spins; $h^{\mu\nu i}_{\rho\sigma j}$ and $\epsilon_{\mu\nu i}$ are scalars; and $s_{\mu\nu i}\coloneqq\op{\mu}{\nu}_i$ is a transition operator for spin $i$.
-If interactions are SU($n$) symmetric then $h^{\mu\nu i}_{\rho\sigma k} = h_{ik} \delta_{\mu\sigma}\delta_{\nu\rho}$, in which case
-\begin{align}
-  H = \f1N \sum_{i<j}
-  h_{ij} \Pi_{ij}
-  + \sum_{\mu,\nu,i} \epsilon_{\mu\nu i} s_{\mu\nu i},
-  &&
-  \Pi_{ij} \equiv \sum_{\mu,\nu} s_{\mu\nu i} s_{\nu\mu j}.
-  \label{eq:spin_sun}
-\end{align}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
-\section{Schwinger boson representation}
-
-We can write the Hamiltonian in \eqref{eq:spin} using Schwinger bosons as
-\begin{align}
-  H = \f1N \sum_{\substack{\mu,\nu,\rho,\sigma\\i<j}}
-  h^{\mu\nu i}_{\rho\sigma j}
-  b_{\mu i}^\dag b_{\nu i} b_{\rho j}^\dag b_{\sigma j}
-  + \sum_{\mu,\nu,i} \epsilon_{\mu\nu i} b_{\mu i}^\dag b_{\nu i},
-\end{align}
-where $b_{\mu i}$ annihilates a boson of type $\mu$ on site $i$.
-The Heisenberg equations of motion for these operators are\footnote{The Hamiltonian in \eqref{eq:spin} only defines $h^{\mu\nu j}_{\rho\sigma k}$ for $j<k$.
-  To simplify expressions in the remainder of this work, we therefore additionally define $h^{\mu\nu k}_{\rho\sigma j}=h^{\rho\sigma j}_{\mu\nu k}$ and $h^{\mu\nu i}_{\rho\sigma i}=0$.}
-\begin{align}
-  \i \d_t b_{\alpha i} = \sp{b_{\alpha i}, H}
-  &= \f1N \sum_{\substack{\mu,\nu,\rho,\sigma\\j<k}}
-  h^{\mu\nu j}_{\rho\sigma k}
-  \sp{b_{\alpha i}, b_{\mu j}^\dag b_{\nu j} b_{\rho k}^\dag b_{\sigma k}}
-  + \sum_{\mu,\nu,j} \epsilon_{\mu\nu j}
-  \sp{b_{\alpha i}, b_{\mu j}^\dag b_{\nu j}} \\
-  &= \f1N \sum_{\mu,\nu,\rho,\sigma,k}
-  h^{\mu\nu j}_{\rho\sigma k}
-  \sp{b_{\alpha i}, b_{\mu i}^\dag b_{\nu i}} b_{\rho k}^\dag b_{\sigma k}
-  + \sum_{\mu,\nu} \epsilon_{\mu\nu i}
-  \sp{b_{\alpha i}, b_{\mu i}^\dag b_{\nu i}} \\
-  &= \sum_{\mu,\nu} \p{\f1N \sum_{\rho,\sigma,k}
-    h^{\mu\nu i}_{\rho\sigma k} b_{\rho k}^\dag b_{\sigma k}
-    + \epsilon_{\mu\nu i}}
-  \sp{b_{\alpha i}, b_{\mu i}^\dag b_{\nu i}}
-\end{align}
-where
-\begin{align}
-  \sp{b_{\alpha i}, b_{\mu i}^\dag b_{\nu i}}
-  = \delta_{\mu\alpha} \delta_{\nu\alpha} b_{\alpha i}
-  + \delta_{\mu\alpha} \p{1-\delta_{\nu\alpha}} b_{\nu i}
-  = \delta_{\mu\alpha} b_{\nu i},
-\end{align}
-so
-\begin{align}
-  \i \d_t b_{\alpha i}
-  = \sum_\nu \p{\f1N \sum_{\rho,\sigma,k}
-    h^{\alpha\nu i}_{\rho\sigma k} b_{\rho k}^\dag b_{\sigma k}
-    + \epsilon_{\alpha\nu i}} b_{\nu i}.
-\end{align}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Interacting regime of a realistic 1D model}
-
-We now consider a realistic 1D model with interactions, in which the Hamiltonian is
-\begin{align}
-  H = -\f{U}{N} \v S^\dag\cdot \v S
-  - J \sum_{\mu,q} \cos\p{q + \mu\varphi} s_{\mu\mu q}.
-\end{align}
-This Hamiltonian can be written in the simplified form
-\begin{align}
-  H = -\f{U}{N} \v S^\dag\cdot \v S
-  - J \sum_q \p{\cos q\, w_{\varphi,q}^+ - \sin q\, w_{\varphi,q}^-},
-\end{align}
-where the diagonal single-spin operators
-\begin{align}
-  w_\varphi^+ \equiv \sum_\mu \cos\p{\mu\varphi} \op{\mu},
-  &&
-  w_\varphi^- \equiv \sum_\mu \sin\p{\mu\varphi} \op{\mu},
-\end{align}
-are respectively even ($+$) and odd ($-$) under spin inversion, $\mu\to-\mu$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Nuclear spin tomography}
-
-Here we discuss the task of performing full tomography on the average reduced single-particle density operator of a collection of nuclear spins.
-As this task will be performed using collective measurements and homogeneous control fields, for ease of language we will consider the equivalent problem of performing tomography on a single nuclear spin.
-We have essentially two ingredients at our disposal: projective measurements of spin onto a fixed quantization axis, and a three-laser drive that addresses nuclear spins via off-resonant coupling to an excited electronic state.
-To be concrete, we can directly measure projectors $\op{\mu}$ onto states of definite spin projection $\mu$ onto the $z$ axis.
-While the three-laser drive gives us access to a variety of nuclear spin Hamiltonians, most notably it allows us to implement arbitrary SU(2) rotations of the form $e^{-\i\v\theta\c\v S}$, with $\v S\equiv\p{S_\x,S_\y,S_\z}$ a vector of spin operators and $\v\theta\equiv\p{\theta_\x,\theta_\y,\theta_\z}$ an arbitrary rotation vector.
-These rotations essentially allow us to measure projectors $\Pi_{\v v\mu} \equiv \op{\mu_{\v v}}$ onto states of definite spin projection $\mu$ onto an arbitrary quantization axis $\v v\in\SS_2$.
-
-Question: given a collection of projectors $\Pi\equiv\set{\Pi_j}$ and measurement outcomes $M\equiv\set{M_j}$, with $M_j$ an empirical estimate of $\bk{\Pi_j}_\rho=\tr\p{\rho\Pi_j}$, how do we determine whether the measurement data $M$ is sufficient to reconstruct $\rho$?
-Answer: reconstruction of $\rho$ is possible if $\Pi$ spans the entire space $\B_n$ of operators on the Hilbert space of an $n$-level spin.
-Given an orthonormal basis $\set{Q_\alpha}$ of self-adjoint operators spanning $\B_n$, we can expand
-\begin{align}
-  \rho = \sum_\alpha \rho_\alpha Q_\alpha,
-  &&
-  \rho_\alpha \equiv \bk{Q_\alpha}_\rho.
-\end{align}
-If $\Pi$ spans $\B_n$, then we can find a set of real numbers $c_{\alpha j}\in\RR$ for which
-\begin{align}
-  Q_\alpha = \sum_j c_{\alpha j} \Pi_j.
-\end{align}
-An empirical estimate $\tilde\rho$ of $\rho$ is then\footnote{This estimate can be improved by defining $\tilde\rho$ as the least-squares fit to the set of linear equations $\tr\p{\tilde\rho\,\Pi_j} = M_j$, and performing maximum-likelihood corrections to $\tilde\rho$ \cite{smolin2012efficient}.}
-\begin{align}
-  \tilde\rho = \sum_\alpha \tilde\rho_\alpha Q_\alpha,
-  &&
-  \tilde\rho_\alpha \equiv \sum_j c_{\alpha j} M_j
-  \approx \sum_j c_{\alpha j} \bk{\Pi_j}_\rho
-  = \bk{Q_\alpha}_\rho
-  = \rho_\alpha.
-\end{align}
-As it turns out, numerical experiments with $n\le20$ suggest that in order to perfurm full tomography of $\rho$, it suffices to measure all projectors $\Pi_{\v v\mu}$ onto states of definite spin $\mu$ along $2n-1$ different axes $\v v$.
-The question remains: which axes should we choose?
-As a first pass, we can simply choose a set of $2n-1$ random axes, $V\equiv\set{\v v}$, by uniformly sampling points on the sphere.
-These random axes define a set of $\p{2n-1}\times n$ projectors $\Pi_V=\set{\Pi^V_j}$, where for shorthand we use a combined index $j=\p{\v v_j,\mu_j}$ that specifies both a measurement axis $\v v_j$ and an outcome $\mu_j$, and for clarity we index objects by $V$ to emphasize their dependence on the choice of measurement axes.
-The projectors $\Pi_V$ can be used to construct $n^2$ linearly independent operators by constructing a matrix $B_V$ with entries
-\begin{align}
-  \sp{B_V}_{ij} = \tr\p{\Pi^V_i \Pi^V_j}.
-\end{align}
-Indexing the (normalized) eigenvectors and the corresponding eigenvalues of $B_V$ by $\alpha$, respectively $\v x^V_\alpha$ and $m^V_\alpha$, for all $\alpha$ with eigenvalue $m^V_\alpha\ne0$ we can then construct the normalized and linearly independent operators
-\begin{align}
-  Q^V_\alpha \equiv \f1{\sqrt{m^V_\alpha}} \sum_j x^V_{\alpha j} \Pi^V_j.
-\end{align}
-To summarize: randomly choosing a set $V$ of $2n-1$ axes will determine a set of measurement operators $\Pi_V$ from which we can construct a complete basis $\set{Q^V_\alpha}$ of operators for $\B_n$.
-However, some choices of axes will be better than others: a good choice of axes $V$ will maximize the amount of information that the corresponding measurement data $M_V=\set{M^V_j}$ will contain about the expectation values $\bk{Q^V_\alpha}_\rho$ used to reconstruct $\rho$.
-The amount of information that $M_V$ contains about $\bk{Q^V_\alpha}_\rho$ is determined, in part, by the magnitude of the eigenvalue $m^V_\alpha$: if $m^V_\alpha$ is small, then $M_V$ will generally contain little information about $\bk{Q^V_\alpha}_\rho$, and vice versa\footnote{Think about the limit $m^V_\alpha\to0$, in which case the fact that $Q^V_\alpha$ is normalized implies that the operator $\sum_j x^V_{\alpha j} \Pi^V_j$ must have a vanishing norm.
-  A vanishing norm can only occur due to fine-tuned cancellations that make the expectation value of $\sum_j x^V_{\alpha j} \Pi^V_j$ difficult to estimate from noisy data on $\bk{\Pi^V_j}_\rho$.}.
-Rather than choosing axes randomly, we can therefore choose a set of axes that minimizes some cost function $C\p{m_V}$ favoring large eigenvalues $m_V\equiv\set{m^V_\alpha}$.
-A reasonable cost function that treats all eigenvalues on equal footing, for example, is
-\begin{align}
-  C_{\t{inv}}\p{m_V} \equiv \sum_\alpha \f1{m^V_\alpha},
-\end{align}
-where the sum is implicitly performed over the $n^2$ largest eigenvalues of $B_V$ (as all other eigenvalues are guaranteed to be zero).
-Note that any information about $\rho$, obtained either from prior knowledge or preliminary measurement data, can be used to construct an even better choice of measurement axes, leading to the possibility of adaptive measurement protocols \cite{pereira2018adaptive} that are more efficient in terms of the number of measurements required to estimate $\rho$ to a fixed precision.
-We leave the details of tailored and adaptive measurement protocols to future work.
-
 \bibliography{notes.bib}

### 2020-09-18 sun_phases/sun_dynamics_notes.tex
moved setspace package before hyperref to fix footnote hyperlinks
fix minor error in notes on raman drive

@@
 \documentclass[nofootinbib,notitlepage,11pt]{revtex4-2}
+
+\usepackage{setspace} % to change text spacing
 
@@
 \cfoot{\thepage~of \pageref{LastPage}} % "x of y" page labeling
-
-\usepackage{setspace} % to change text spacing
 
@@
   \p{e^{\i\phi_m j} s_{mj} \otimes\op{\up}{\dn}_j + \t{h.c.}}
-  + \f{\Delta}{2}\, \1 \otimes\p{\op\up_j-\op\dn_j},
+  + \f{\Delta}{2} \sum_j \1_j \otimes\p{\op\up_j-\op\dn_j},
   \label{eq:drive_full}
 \end{align}
-where $j$ indexes a lattice site; $s_\pm$ are spin-raising ($+$) and spin-lowering ($-$) operators (containing all appropriate Clebsch-Gordan coefficients); $s_0\coloneqq s_\z$ is a spin-$z$ operator; $\Delta$ is the detuning of the drive lasers from the excited-state transition frequency; and $\1$ is an identity operator (acting on the nuclear spins of all atoms).
+where $j$ indexes a lattice site; $s_\pm$ are spin-raising ($+$) and spin-lowering ($-$) operators (containing all appropriate Clebsch-Gordan coefficients); $s_0\coloneqq s_\z$ is a spin-$z$ operator; and $\1$ is an identity operator on an $n$-level spin.
 

### 2020-09-22 collective_sun/sun_perturbation.tex
fixed minor error

@@
 orthonormal basis
-$\col\B_{r_{\t{max}}}\equiv\bigotimes_{r\le r_{\t{max}}}\B_r$ for the
+$\col\B_{r_{\t{max}}}\equiv\bigoplus_{r\le r_{\t{max}}}\B_r$ for the
 combined manifold of all states with at most $r_{\t{max}}$ magnons.

### 2020-09-30 sun_phases/sun_dynamics_notes.tex
properly account for laser polarization in nuclear spin drive

@@
 
+\newcommand{\CC}{\mathbb{C}}
 \newcommand{\RR}{\mathbb{R}}
@@
 \newcommand{\ZZ}{\mathbb{Z}}
+
+\newcommand{\su}{\mathfrak{su}}
 
@@
 We will also consider two types of initial states: a fully polarized state $\ket\X$ in which all spins point along the $x$ axis, and a ``back-to-back'' state $\ket\BTBX$ in which half of the spins are polarized along $+x$, and half along $-x$.
-Specifically, indexing spins by their associated quasi-momentum mode $q\in[-\pi,\pi)$, the initial state $\ket\BTBX$ has spins with $q\ge0$ ($q<0$) initially polarized along $+x$ ($-x$).
+Specifically, indexing spins by their associated quasi-momentum mode $q\in(-\pi,\pi)$, the initial state $\ket\BTBX$ has spins with $q>0$ ($q<0$) initially polarized along $+x$ ($-x$).
 In this way, the initial state $\ket\BTBX$ is correlated with $\sim\sin q$ terms of the spin Hamiltonian in \eqref{eq:spin_pair}.
@@
 
-Figures \ref{fig:ss_spect_n2_X_p0.5}--\ref{fig:ss_spect_n2_XX_p1.0} show the time-series of $s^2(t)$, its power spectral density $\widetilde{s^2}(\omega)$, and time-series spectrum of $\bar\rho(t)$ for the case of $n=2$, and with values of $J/U$ between $10^{-2}$ and $10^1$.
+Figures \ref{fig:ss_spect_n2_X_p0.5}--\ref{fig:ss_spect_n2_XX_p1.0} show the time-series of $s^2(t)$, its power spectral density $\widetilde{s^2}(\omega)$, and the time-series spectrum of $\bar\rho(t)$ for the case of $n=2$, with values of $J/U$ between $10^{-2}$ and $10^1$.
 In short, these figures all tell the same story: there is a ``collective'' phase when $J/U\lesssim1$, and of a ``single-particle'' phase when $J/U\gtrsim1$.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Nuclear spin tomography}
+\section{Spin qudit tomography}
 \label{sec:tomography}
+
+% Motivations / background:
+% https://arxiv.org/pdf/1901.06282.pdf
+% https://arxiv.org/abs/1911.00099
+% https://arxiv.org/abs/2005.10910
+% https://journals.aps.org/pra/abstract/10.1103/PhysRevA.86.012118
+% https://journals.aps.org/pra/abstract/10.1103/PhysRevA.94.010301
+% https://arxiv.org/pdf/2003.08668.pdf
+% https://www.osapublishing.org/ol/abstract.cfm?uri=ol-44-10-2558
 
@@
 In addition to letting atoms evolve freely under $H_{\t{spin}}$, we can turn on external control fields to address atoms' internal states.
-Our primary tool here will be three lasers that off-resonantly address an excited electronic state of the atoms.
-Each laser will be indexed by a different integer $m\in\set{1,0,-1}$ that determines the nuclear spin transition $\mu\to\mu+m$ accompanying the electronic $\ket\dn\to\ket\up$ transition upon absorption of a photon.
-Laser $m$ will have a homogeneous drive amplitude $\Omega_m$, and a relative phase $e^{\i\phi_m}$ between adjacent lattice sites.
-All lasers will be detuned by a frequency $\Delta$ away from the electronic $\ket\dn\to\ket\up$ transition.
+Our primary tool here will be a multi-laser drive that off-resonantly addresses an excited electronic state of the atoms.
+Specifically, all lasers will have a frequency that is detuned by $\Delta$ from an electronic $\ket\dn\to\ket\up$ transition.
+Each laser will be indexed by an axis of propagation, $\v v$, with right/left-circular polarization amplitudes $\Omega_{\v v\pm} e^{\i\eta_{\v v\pm}}$ (with real $\Omega_{\v v\pm}$ and $\eta_{\v v\pm}$).
+An electronic $\ket\dn\to\ket\up$ transition is thus accompanied by a nuclear spin transition $\mu_{\v v}\to\mu_{\v v}\pm1$, where $\mu_{\v v}$ is the projection of nuclear spin onto $\v v$.
+Furthermore, laser $\v v$ will generally have a relative phase $e^{\i\kappa\v v\c\v\ell}$ between neighboring lattice sites, where $\kappa$ is the wavenumber of the drive lasers (in units of the lattice spacing) and $\v\ell$ is a fixed unit vector parallel to the lattice.
 Altogether, these lasers induce the drive Hamiltonian
 \begin{align}
-  H_{\t{drive}}^{\t{full}} = \sum_j \sum_{m\in\set{1,0,-1}} \Omega_m
-  \p{e^{\i\phi_m j} s_{mj} \otimes\op{\up}{\dn}_j + \t{h.c.}}
+  H_{\t{drive}}^{\t{full}} = \sum_{j,\v v,\sigma}
+  \Omega_{\v v\sigma} \p{e^{\i\eta_{\v v\sigma} + \i\kappa\v v\c\v\ell j}
+    s_{\v v\sigma j} \otimes \op{\up}{\dn}_j + \t{h.c.}}
   + \f{\Delta}{2} \sum_j \1_j \otimes\p{\op\up_j-\op\dn_j},
-  \label{eq:drive_full}
 \end{align}
-where $j$ indexes a lattice site; $s_\pm$ are spin-raising ($+$) and spin-lowering ($-$) operators (containing all appropriate Clebsch-Gordan coefficients); $s_0\coloneqq s_\z$ is a spin-$z$ operator; and $\1$ is an identity operator on an $n$-level spin.
-
+where $s_{\v v\pm}$ are spin-raising/lowering operators along axis $\v v$ (containing all appropriate Clebsch-Gordan coefficients), and $\1$ is the identity operator for a single $n$-level spin (with total spin $I=\frac{n-1}{2}$).
+To write out $s_{\v v\pm}$ in full, we first identify the operators $s_\x,s_\y,s_\z$ that generate rotations of an $n$-level spin in 3D space, and define $s_\pm\equiv s_\x\pm\i s_\y$.
+Denoting the polar and azimuthal angles of $\v v$ respectively by $\alpha_{\v v}$ and $\beta_{\v v}$, we can use the fact that $s_\x,s_\y,s_\z$ form an $\su(2)$ algebra to expand
+\begin{align}
+  s_{\v v\pm}
+  &= e^{-\i\beta_{\v v}s_\z} e^{-\i\alpha_{\v v}s_\y}
+  s_\pm e^{\i\alpha_{\v v}s_\y} e^{\i\beta_{\v v}s_\z} \\
+  &= -\sin\alpha_{\v v}\, s_\z
+  + \f12 \p{\cos\alpha_{\v v}+1} e^{\mp\i\beta_{\v v}} s_\pm
+  + \f12 \p{\cos\alpha_{\v v}-1} e^{\pm\i\beta_{\v v}} s_\mp.
+\end{align}
+We now consider a particular setup involving three drive lasers: one linearly polarized laser oriented along the $x$ axis, with propagation axis $\v v_0\equiv\p{1,0,0}$ and polarization amplitudes $\Omega_{\v v_\x\sigma}e^{\i\eta_{\v_\x\sigma}}=-\Omega_0e^{\i\eta_0}/2$ (for both $\sigma\in\set{\pm1}$); two right-circularly polarized lasers pointing in opposite directions along the $z$ axis, with propagation axes $\v v_\pm\equiv\p{0,0,\pm1}$ and polarization amplitudes $\Omega_{\v v_\pm,+}e^{\i\eta_{\v_\pm,+}}=\pm\Omega_\pm e^{\i\eta_\pm}$; and a 1D lattice oriented in the $y$-$z$ plane at an angle $\theta$ from the $z$ axis, with lattice axis $\v\ell=\p{0,\sin\theta,\cos\theta}$ [\todo{make figure?}].
+In this setup, the relevant spin-raising/lowering operators are
+\begin{align}
+  s_{\v v_0\pm} = -s_\z \pm \i s_\y,
+  &&
+  s_{\v v_\pm,+} = \pm s_\pm,
+\end{align}
+and the site-dependent phases are $e^{\i\kappa\v v_m\c\v\ell j}=e^{\i m\phi j}$, where $m\in\set{0,+1,-1}$ and we define $\phi\equiv\kappa\cos\theta$ for shorthand.
+Altogether, the Hamiltonian induced by this three laser drive is
+\begin{align}
+  H_{\t{drive}}^{\t{3LD}} = \sum_{j,m} \Omega_m
+  \p{e^{\i\eta_m+\i m\phi j} s_{mj} \otimes\op{\up}{\dn}_j + \t{h.c.}}
+  + \f{\Delta}{2} \sum_j \1_j \otimes\p{\op\up_j-\op\dn_j},
+\end{align}
+where we define $s_0\equiv s_\z$ for convenience of notation.
 In the far-detuned limit $\Delta\gg\Omega_m$, a perturbative treatment of the excited electronic state yields an effective drive Hamiltonian that only addresses nuclear spins.
-Furthermore, if the angles $\phi_m$ satisfy $\phi_m-\phi_{m'}=\p{m-m'}\phi$ for a fixed spin-orbit coupling angle $\phi$ (e.g.~with $\phi_m=m\phi$), then the gauge transformation $c_{j\mu}\to e^{\i\mu\phi j} c_{j\mu}$ makes the drive Hamiltonian spatially homogeneous.
-Altogether, we arrive at a homogeneous effective drive Hamiltonian of the form
+Furthermore, the gauge transformation $c_{j\mu}\to e^{\i\mu\phi j} c_{j\mu}$ makes this effective Hamiltonian spatially homogeneous, taking the form
 \begin{align}
-  H_{\t{drive}} = \sum_j H_{\t{drive},j}^{\t{single}},
+  H_{\t{drive}}^{\t{eff}} = \sum_j H_{\t{drive},j}^{\t{single}},
 \end{align}
-where the single-spin Hamiltonian $H_{\t{drive}}^{\t{single}}$ can be conveniently expressed in terms of operators $s_\x,s_\y,s_\z$ that generate rotations of an $n$-level spin (with total spin $I=\frac{n-1}{2}$) in 3D space:
+where the single-spin Hamiltonian $H_{\t{drive}}^{\t{single}}$ is given by
+\begin{multline}
+  \Delta \times H_{\t{drive}}^{\t{single}}
+  = \p{\Omega_+^2 - \Omega_-^2} s_\z
+  - \sum_{\sigma\in\set{\pm1}} \Omega_0 \Omega_\sigma
+  \p{\sigma s_{\tilde\eta_\sigma,\x}
+    + s_\z s_{\tilde\eta_\sigma,\x} + s_{\tilde\eta_\sigma,\x} s_\z} \\
+  - \Omega_0^2 s_\z^2
+  - \p{\Omega_+ + \Omega_-}^2 s_{\tilde\eta_0,\x}^2
+  - \p{\Omega_+ - \Omega_-}^2 s_{\tilde\eta_0,\y}^2
+  ,
+  \label{eq:drive_single}
+\end{multline}
+with
 \begin{align}
-  H_{\t{drive}}^{\t{single}}
-  = h_+ h_- s_\z + h_0 h_+ s_\x
-  + h_0 h_- \p{s_\z s_\x  + s_\x s_\z}
-  - h_0^2\p{s_\z^2-\v s_{\t{3D}}^2} - h_-^2 s_\x^2 - h_+^2 s_\y^2.
+  \tilde\eta_\pm \equiv \pm\p{\eta_0 - \eta_\pm},
+  &&
+  \tilde\eta_0 \equiv \f{\tilde\eta_+ - \tilde\eta_-}{2}
+  = -\f{\eta_+ - \eta_-}{2},
+  &&
+  s_{\varphi m} = e^{-\i\varphi s_\z} s_m e^{\i\varphi s_\z}.
 \end{align}
-with $\v s_{\t{3D}}\equiv\p{s_\x,s_\y,s_\z}$.
-Here $h_0\equiv\Omega_0/\sqrt\Delta$ and $h_\pm\equiv\p{\Omega_+\pm\Omega_-}/\sqrt{2\Delta}$ can be tuned independently through the choice of drive amplitudes $\Omega_m$; and $\v s_{\t{3D}}^2=s_\x^2+s_\y^2+s_\z^2=I\p{I+1}$ is a constant.
-Table \ref{tab:drives} shows several drive Hamiltonians that can be implemented with various amplitude-matching conditions.
-We do not expect these controls to be sufficient for implementing arbitrary collective SU($n$) rotations, but identifying the subgroup of SU($n$) generated by these controls could be an interesting problem for future work.
-One notable capability is using pulse sequences to implement SU(2) rotations of the form $e^{-\i\v\theta\c\v s_{\t{3D}}}$ with an arbitrary rotation vector $\v\theta\equiv\p{\theta_\x,\theta_\y,\theta_\z}$.
-In addition to enabling us to perform arbitrary 3D rotations, these collective rotations enable us to measure all components of the mean collective spin matrix $\bk{\v S}$, as discussed in Appendix \ref{sec:tomography}.
+We do not expect these controls to be sufficient for implementing arbitrary collective SU($n$) rotations, but identifying the subgroup of SU($n$) that can be generated by \eqref{eq:drive_single} could be an interesting problem for future work.
+A simple special case to consider is that of all $\eta_m=0$, in which case
+\begin{align}
+  \Delta \times H_{\t{drive}}^{\t{single}}
+  = \tilde\Omega_+ \tilde\Omega_- s_\z
+  + \tilde\Omega_0 \tilde\Omega_- s_\x
+  + \tilde\Omega_0 \tilde\Omega_+ \p{s_\z s_\x  + s_\x s_\z}
+  - \tilde\Omega_0^2 s_\z^2 - \tilde\Omega_+^2 s_\x^2
+  - \tilde\Omega_-^2 s_\y^2,
+\end{align}
+where we define $\tilde\Omega_0=-\Omega_0$ and $\tilde\Omega_\pm\equiv\Omega_+\pm\Omega_-$ for convenience.
+Table \ref{tab:drives} shows several drive Hamiltonians that can be implemented with various amplitude-matching conditions (choices of $\tilde\Omega_m$) when all $\eta_m=0$.
+One notable capability that these controls provide us is that of using pulse sequences to implement SU(2) rotations of the form $e^{-\i\v\theta\c\v s_{\t{3D}}}$, where $\v s_{\t{3D}}\equiv\p{s_\x,s_\y,s_\z}$ and $\v\theta\equiv\p{\theta_\x,\theta_\y,\theta_\z}$ is an arbitrary rotation vector.
+In addition to enabling arbitrary 3D rotations, these collective rotations enable us to measure all components of the mean collective spin matrix $\bk{\v S}$, as discussed in Appendix \ref{sec:tomography}.
 
@@
   \centering
-  \caption{Drive Hamiltonians that can be implemented with various amplitude-matching conditions.}
-  \begin{tabular}{c|c|c}
-    Condition & $\p{h_0,h_+,h_-}$
+  \caption{Drive Hamiltonians that can be implemented with various amplitude-matching conditions when all $\eta_m=0$.
+    Here $\v s_{\t{3D}}^2=s_\x^2+s_\y^2+s_\z^2=I\p{I+1}$ is a constant.}
+  \begin{tabular}{c|c}
+    $(\tilde\Omega_0,\tilde\Omega_+,\tilde\Omega_-)/\sqrt{\Delta}$
     & $H_{\t{drive}}^{\t{single}}$
     \\ \hline\hline
-    $\Omega_+=\Omega_-=0$ & $\p{1,0,0}$ & $-s_\z^2 + \v s_{\t{3D}}^2$
+    $\p{1,0,0}$ & $-s_\z^2$
     \\ \hline
-    $\Omega_0=0$,  $\Omega_+=+\Omega_-$ & $\p{0,1,0}$ & $-s_\y^2$
+    $\p{0,1,0}$ & $-s_\x^2$
     \\ \hline
-    $\Omega_0=0$,  $\Omega_+=-\Omega_-$ & $\p{0,0,1}$ & $-s_\x^2$
+    $\p{0,0,1}$ & $-s_\y^2$
     \\ \hline
-    $\Omega_0=\Omega_\mp=0$ & $\p{0,1,\pm1}$
-    & $\pm s_\z + s_\z^2 - \v s_{\t{3D}}^2$
+    $\p{0,1,\pm1}$ & $\pm s_\z + s_\z^2 - \v s_{\t{3D}}^2$
     \\ \hline
-    $\sqrt{2}\Omega_0=\pm\Omega_+=\pm\Omega_-$
-    & $\p{1,\pm1,0}$ & $\pm s_\x + s_\x^2$
+    $\p{1,0,\pm1}$ & $\pm s_\x + s_\x^2 - \v s_{\t{3D}}^2$
     \\ \hline
-    $\sqrt{2}\Omega_0=\pm\Omega_+=\mp\Omega_-$
-    & $\p{1,0,\pm1}$ & $\pm\p{s_\z s_\x+s_\x s_\z} + s_\y^2$
-    \\ \hline
-    $\Omega_0=\pm\Omega_+=\pm\Omega_-$ & $\p{1,\pm\sqrt{2},0}$
-    & $\pm\sqrt{2} s_\x + \p{s_\x^2 - s_\y^2}$
-    \\ \hline
-    $\Omega_0=\pm\Omega_+=\mp\Omega_-$ & $\p{1,0,\pm\sqrt{2}}$
-    & $\pm\sqrt{2}\p{s_\z s_\x + s_\x s_\z} - \p{s_\x^2-s_\y^2}$
+    $\p{1,\pm1,0}$
+    & $\pm\p{s_\z s_\x+s_\x s_\z} + s_\y^2 - \v s_{\t{3D}}^2$
   \end{tabular}
@@
 The Heisenberg equations of motion for these operators are\footnote{The Hamiltonian in \eqref{eq:spin} only defines $h^{\mu\nu j}_{\rho\sigma k}$ for $j<k$.
-  To simplify expressions in the remainder of this work, we therefore additionally define $h^{\rho\sigma k}_{\mu\nu j}=h^{\mu\nu j}_{\rho\sigma k}$ and $h^{\mu\nu j}_{\rho\sigma j}=0$.}
+  To simplify expressions, we therefore additionally define $h^{\rho\sigma k}_{\mu\nu j}=h^{\mu\nu j}_{\rho\sigma k}$ and $h^{\mu\nu j}_{\rho\sigma j}=0$.}
 \begin{align}

### 2020-10-14 sun_phases/spin_qudit_tomo.tex
add draft/notes for spin qudit tomography paper

\documentclass[nofootinbib,notitlepage,twocolumn]{revtex4-2}

\usepackage{setspace} % to change text spacing

%%% linking references
\usepackage{hyperref}
\hypersetup{
  breaklinks=true,
  colorlinks=true,
  linkcolor=blue,
  filecolor=magenta,
  urlcolor=cyan,
}

%%% header / footer
\usepackage{fancyhdr} % easier header and footer management
\pagestyle{fancy} % page formatting style
\fancyhf{} % clear all header and footer text
\renewcommand{\headrulewidth}{0pt} % remove horizontal line in header
\usepackage{lastpage} % for referencing last page
\cfoot{\thepage~of \pageref{LastPage}} % "x of y" page labeling

%%% symbols, notations, etc.
\usepackage{physics,braket,bm,amssymb} % physics and math
\renewcommand{\t}{\text} % text in math mode
\newcommand{\f}[2]{\dfrac{#1}{#2}} % shorthand for fractions
\newcommand{\p}[1]{\left(#1\right)} % parenthesis
\renewcommand{\sp}[1]{\left[#1\right]} % square parenthesis
\renewcommand{\set}[1]{\left\{#1\right\}} % curly parenthesis
\newcommand{\bk}{\Braket} % shorthand for braket notation
\renewcommand{\v}{\bm} % bold vectors
\newcommand{\uv}[1]{\bm{\hat{#1}}} % unit vectors
\renewcommand{\c}{\cdot} % inner product
\renewcommand{\i}{\mathrm{i}\mkern1mu} % imaginary unit

\usepackage{dsfont} % for identity operator
\newcommand{\1}{\mathds{1}}

\usepackage{mathtools} % for \coloneqq

\newcommand{\up}{\uparrow}
\newcommand{\dn}{\downarrow}
\newcommand{\x}{\text{x}}
\newcommand{\y}{\text{y}}
\newcommand{\z}{\text{z}}

\newcommand{\B}{\mathcal{B}}
\newcommand{\G}{\mathcal{G}}
\renewcommand{\O}{\mathcal{O}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\T}{\mathcal{T}}
\newcommand{\Y}{\mathcal{Y}}

\newcommand{\D}{\mathfrak{D}}
\renewcommand{\dd}{\mathfrak{d}}
\renewcommand{\d}{\text{d}}

\newcommand{\SO}{\text{SO}}

\newcommand{\CC}{\mathbb{C}}
\newcommand{\RR}{\mathbb{R}}
\renewcommand{\SS}{\mathbb{S}}
\newcommand{\ZZ}{\mathbb{Z}}

\newtheorem{theorem}{Theorem}[section]

\def\obra#1{\mathinner{({#1}|}}
\def\oket#1{\mathinner{|{#1})}}
\def\obk#1{\mathinner{({#1})}}
\def\oop#1#2{\oket{#1}\!\obra{#2}}

%%% figures
\usepackage{graphicx} % for figures
\graphicspath{{./figures/}} % set path for all figures

% to place figures in the correct section
\usepackage[section]{placeins}

\usepackage[inline]{enumitem} % in-line lists and \setlist{} (below)
\setlist[enumerate,1]{label={(\roman*)}} % default in-line numbering
\setlist{nolistsep} % more compact spacing between environments

%%% text markup
\usepackage{color} % text color
\newcommand{\red}[1]{{\color{red} #1}}
\newcommand{\todo}[1]{{\color{blue} To do: #1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\newcommand{\JILA}{JILA, National Institute of Standards and Technology and
  University of Colorado, 440 UCB, Boulder, Colorado 80309, USA}
\newcommand{\CTQM}{Center for Theory of Quantum Matter, University of Colorado, Boulder, CO, 80309, USA}

\newcommand{\thetitle}{Spin qudit tomography}

\title{\thetitle}
\author{Michael A.~Perlin}
\email{mika.perlin@gmail.com}
\author{Ana Maria Rey}
\affiliation{\JILA}
\affiliation{\CTQM}
\date{\today}

\keywords{spin qudits; quantum state tomography}

\begin{abstract}
  We consider the task of performing quantum state tomography on a $d$-state spin qudit, using only measurements of spin projection onto different quantization axes.
  By mapping this quantum tomography problem onto the well-studied classical problem of signal recovery on the sphere, we prove that full reconstruction of arbitrary mixed states is possible, and requires a minimum number of measurement axes, $r_d$, that is bounded by $2d-1\le r_d\le d^2$.
  We then provide an explicit optimized tomographic protocol, as well as numerical experiments to show that $r_d=2d-1$ for $d\le20$.
  [\red{Note: there are also a few other things that I want to look into, such as (1) adaptive tomography protocols, (2) proofs about something like the number of measurements necessary to reconstruct an arbitrary state to a given precision, (3) a simplified tomography procedure for measuring pure states, (4) the trade-off between measurement precision and the \# of measurement axes (for a fixed total number of measurements).]}
\end{abstract}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

Quantum state tomography, or the task of reconstructing a quantum state by collecting and processing measurement data, is an essential algorithmic primitive for quantum sensing, quantum simulation, and quantum information processing.
The central importance of quantum state tomography has led to the development of a variety of techniques based on compressed sensing \cite{gross2010quantum}, matrix product states \cite{cramer2010efficient}, % confidence regions \cite{christandl2012reliable},
maximum-likelihood estimation \cite{smolin2012efficient}, % linear \cite{qi2013quantum} and least-squares \cite{opatrny1997leastsquares} regression,
Bayesian inference \cite{huszar2012adaptive}, and neural networks \cite{torlai2018neuralnetwork}, among others that are too numerous to list here.
These techniques are typically developed in a general, information-theoretic setting, and make minimal assumptions about the physical medium of a quantum state.
As a consequence, even well-established techniques can nonetheless be ill-suited for physical platforms with unique or limited capabilities.

[\red{Discuss spin qudits as quantum resources for quantum sensing \cite{stefano2019set}, quantum simulation \cite{banerjee2013atomic, cazalilla2014ultracold, rico2018nuclear}, and quantum information processing \cite{albert2020robust, gross2020encoding}}]

[\red{Discuss previous work on qudit tomography \cite{thew2002qudit, flammia2005minimal, salazar2012quantum, sosa-martinez2017quantum, ha2018minimal, evrard2019enhanced, stefano2019set, palici2020oam}}]

[\red{Discuss practical limitations with spin qudits: the operations necessary for existing tomographic protocols may be inaccessible, difficult to implement, or otherwise more costly than simple spin projection measurements.}]

[\red{Mention previous work on spin tomography \cite{manko1997spin}.  Discuss connection to quantum state tomography through the discrete Wigner function \cite{leonhardt1995quantumstate, leonhardt1996discrete}.}]

[\red{Discuss connection to signal recovery on the sphere.  Original $\sim4d^2$ signaling theorem \cite{driscoll1994computing}; its $\sim2d^2$ refinement \cite{mcewen2011novel}; a nice review of these methods \cite{mcewen2011sampling}; an empirically accurate (albeit without rigorous proofs/guarantees) $d^2$ sampling method \cite{khalid2014optimaldimensionality}; current status of the theory behind sparse signal recovery with random sampling \cite{rauhut2011sparse} (note: existing bounds are {\it not} believed to be tight); nice numerical method for sparse signal recovery \cite{alem2012sparse}; useful textbook references \cite{freeden2008spherical, freeden2018spherical}.}]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Spin transition operators and rotation matrices}

Consider a $d$-state spin qudit with total spin $s\equiv\frac{d-1}{2}$.
The defining property of a spin qudit is the fact that it describes an angular momentum degree of freedom, which has specific implications for how a spin qudit should transform under the group SO(3) of rotations in 3D space.
Due to the central importance of these transformation rules for a spin qudit, we first identify a basis of operators that transforms nicely under 3D rotations.
One such basis is provided by the {\it transition operators} (also known as polarization operators \cite{kryszewski2006positivity, bertlmann2008bloch}), defined by
\begin{align}
  T_{\ell m} \equiv \sqrt{\f{2\ell+1}{2s+1}} \sum_{\mu,\nu=-s}^s
  \bk{s\mu;\ell m|s\nu} \op{\nu}{\mu},
  \label{eq:trans_op}
\end{align}
where $\bk{s\mu;\ell m|s\nu}$ is a Clebsh-Gordan coefficient that enforces $\ell\in\set{0,1,\cdots,2s}$ and $m\in\set{-\ell,-\ell+1,\cdots,\ell}$, such that there are $d^2$ operators in total.
For brevity, we will generally treat the value of $d$ (or $s$) as constant but arbitrary throughout this work, and suppress any explicit dependence of quantities or operators such as $T_{\ell m}$ on $d$.
The transition operators are orthonormal with respect to the trace inner product $\obk{\O|\Q}\equiv\tr\p{\O^\dag\Q}$, and transform nicely under conjugation:
\begin{align}
  \obk{T_{\ell m }|T_{\ell'm'}}
  = \delta_{\ell\ell'} \delta_{mm'},
  &&
  T_{\ell m}^\dag = \p{-1}^m T_{\ell,-m},
\end{align}
where $\delta_{kk'}\equiv 1$ if $k=k'$ and $0$ otherwise.
As a consequence, we can expand any density operator $\rho$ in the transition operator basis as
\begin{align}
  \rho = \sum_{\ell=0}^{2s} \sum_{m=-\ell}^\ell \rho_{\ell m} T_{\ell m},
  &&
  \rho_{\ell m}\equiv\obk{T_{\ell m}|\rho},
  \label{eq:trans_state}
\end{align}
where $\rho^\dag=\rho$ implies that $\rho_{\ell m}^*=\p{-1}^m\rho_{\ell,-m}$.
The transition operators can be physically interpreted as the transition $\ket\psi\to T_{\ell m}\ket\psi$ induced on the state $\ket\psi$ of a spin-$s$ system by the absorption of a spin-$\ell$ particle with spin projection $m$ onto a fixed quantization axis.
In analogy with the spherical harmonics $Y_{\ell m}$, we will refer to the indices $\ell$ and $m$ respectively as the {\it degree} and {\it order} of $T_{\ell m}$.

The transition operators are spherical tensor operators, whose degree $\ell$ is preserved under 3D rotations generated by the spin operators $S_\x,S_\y,S_\z$ (see Appendix \ref{sec:rotations}).
For any triplet of Euler angles $\v\omega=\p{\alpha_{\v\omega},\beta_{\v\omega},\gamma_{\v\omega}}$, we can therefore define the {\it fundamental} rotation operator
\begin{align}
  R\p{\v\omega} \equiv e^{-\i\alpha_{\v\omega}S_\z} e^{-\i\beta_{\v\omega}S_\y} e^{-\i\gamma_{\v\omega}S_\z},
  \label{eq:rot_op}
\end{align}
as well as the corresponding {\it adjoint} rotation operator $\R\p{\v\omega}$ defined by $\R\p{\v\omega}\O = R\p{\v\omega} \O R\p{\v\omega}^\dag$, and expand
\begin{align}
  T_{\v\omega\ell m} \equiv
  \R\p{\v\omega} T_{\ell m}
  = \sum_{n=-\ell}^\ell \D_{nm}^\ell\p{\v\omega} T_{\ell n}.
  \label{eq:trans_rot}
\end{align}
The adjoint rotation matrix elements $\D_{mn}^\ell\p{\v\omega}\equiv\obk{T_{\ell m}|\R\p{\v\omega}|T_{\ell n}}$ are an analog of the fundamental (Wigner) rotation matrix elements $D_{mn}^\ell\p{\v\omega}\equiv\bk{\ell m|R\p{\v\omega}|\ell n}$ that govern rotations of spin-$\ell$ particles, where $\ket{\ell m}$ denotes the state of a spin-$\ell$ qudit with spin projection $m$ onto a quantization axis.
Throughout this work, we will primarily consider rotations determined by points on the 2-sphere $\Omega_2$ at $\gamma_{\v\omega}=0$.
For ease of notation, we therefore implicitly associate doublets $\v v=\p{\alpha_{\v v},\beta_{\v v}}$ with the triplet $\p{\alpha_{\v v},\beta_{\v v},0}$, so that for example $R\p{\v v} = R\p{\alpha_{\v v},\beta_{\v v},0}$ and $T_{\v v\ell m} = T_{\p{\alpha_{\v v},\beta_{\v v},0},\ell m}$.

As functions of $\v\omega$, the fundamental rotation matrix elements $D_{mn}^\ell$ satisfy the orthogonality relations \cite{brown2003rotational}
\begin{align}
  \int_{\Omega_3} \d\Omega_3\p{\v\omega}
  D^{\ell}_{mn}\p{\v\omega}^* D^{\ell'}_{m'n'}\p{\v\omega}
  = \f{8\pi^2}{2\ell+1} \delta_{\ell\ell'} \delta_{mm'} \delta_{nn'},
  \label{eq:ortho}
\end{align}
where integration is performed over all $\v\omega\in\Omega_3\simeq[0,2\pi]\times[0,\pi]\times[0,2\pi]$ with integration measure $\d\Omega_3\p{\v\omega} \equiv \d\alpha_{\v\omega} \d\beta_{\v\omega} \d\gamma_{\v\omega} \sin\beta_{\v\omega}$.
In particular, the restriction of $D^\ell_{m,0}$ to the 2-sphere $\Omega_2$ at $\gamma_{\v\omega}=0$ essentially recovers (up to normalization factors and complex conjugation) the spherical harmonics $Y_{\ell m}$.
The same orthogonality relations in Eq.~\eqref{eq:ortho} are inherited by the rotation matrix elements $\D_{mn}^\ell$ through the fact $\R\p{\v\omega}$ is the adjoint action of $R\p{\v\omega}$ (see Appendix \ref{sec:ortho}), and the functions $\set{\D_{m,0}^\ell}$ restricted to $\Omega_2$ are similarly an orthogonal basis for the space $L^2\p{\Omega_2}$ of square-integrable functions on the sphere.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Spin tomography as signal recovery on the sphere}

Our goal is to reconstruct an arbitrary state of a spin qudit from measurements of spin projection onto different quantization axes.
We are thus nominally restricted to measuring projectors $\Pi_{\v v\mu} \equiv \op{\mu_{\v v}}$, where $\ket{\mu_{\v v}}\equiv R\p{\v v}\ket{\mu}$ is a state with spin projection $\mu$ onto a measurement axis determined by a point $\v v$ on the 2-sphere $\Omega_2$.
For any fixed axis $\v v$, the sets $\set{\Pi_{\v v\mu}}$ and $\set{T_{\v v\ell,0}}$ are both orthonormal bases for the space of diagonal operators in the basis $\set{\ket{\mu_{\v v}}}$.
By a simple change of operator basis, measuring the projectors $\set{\Pi_{\v v\mu}}$ is therefore equivalent to measuring the transition operators $\set{T_{\v v\ell,0}}$, and provides data on the expectation values $\bk{T_{\v \ell,0}}_\rho\equiv\tr\p{\rho T_{\v v\ell,0}}$.

Reconstructing an arbitrary density operator $\rho$ essentially requires us to find a set of axes $V=\set{\v v}$ and associated coefficients $C_{\ell mk}\p{\v v}$ that would allow us to recover any coefficient $\rho_{\ell m}$ of the expansion in Eq.~\eqref{eq:trans_state} through
\begin{align}
  \rho_{\ell m}^* = \bk{T_{\ell m}}_\rho
  = \sum_{\v v,k} C_{\ell mk}\p{\v v} \bk{T_{\v vk,0}}_\rho.
  \label{eq:state_recon}
\end{align}
Expanding the rotated transition operators $T_{\v vk,0}$ according to Eq.~\eqref{eq:trans_rot}, we find that the recovery condition in Eq.~\eqref{eq:state_recon} is satisfied when
\begin{align}
  T_{\ell m}
  = \sum_{\v v,k,n} C_{\ell mk}\p{\v v} \D^k_{n,0}\p{\v v} T_{kn},
\end{align}
where the orthogonality of the transition operators motivates decomposing $C_{\ell mk}\p{\v v}=\delta_{k\ell}C_{\ell m}\p{\v v}$, which in turn implies that
\begin{align}
  \sum_{\v v} C_{\ell m}\p{\v v} \D^\ell_{n,0}\p{\v v} = \delta_{mn}
  \label{eq:tomo_recovery}
\end{align}
for all $\ell$.
In fact, the problem of finding suitable axes $V$ and coefficients $C_{\ell m}\p{\v v}$ to satisfy Eq.~\eqref{eq:tomo_recovery}, solving which would enable the recovery of arbitrary qudit states through Eqs.~\eqref{eq:trans_state} and \eqref{eq:state_recon}, can be mapped onto the well-studied problem of signal recovery on the sphere \cite{driscoll1994computing, mcewen2011novel, mcewen2011sampling, khalid2014optimaldimensionality, rauhut2011sparse, alem2012sparse}.
The signal recovery problem can be stated as follows: given a square-integrable function $f\in L^2\p{\Omega_2}$ on the sphere, with the harmonic expansion
\begin{align}
  f\p{\v v} = \sum_{\ell,m} f_{\ell m} B_{\ell m}\p{\v v},
\end{align}
where $f_{\ell m}$ are expansion coefficients and $\set{B_{\ell m}}$ is an orthogonal basis for $L^2\p{\Omega_2}$, find a set of points $V=\set{\v v}$ and associated coefficients $C_{\ell m}\p{\v v}$ with which we can reconstruct $f$, or equivalently its coefficients $f_{\ell m}$, from knowledge of the function's value $f\p{\v v}$ at all points $\v v\in V$; that is
\begin{align}
  f_{\ell m} = \sum_{\v v} C_{\ell m}\p{\v v} f\p{\v v}
  = \sum_{\v v,k,n} C_{\ell m}\p{\v v} f_{kn} B_{kn}\p{\v v}.
\end{align}
Reconstruction of functions with arbitrary coefficients $f_{\ell m}$ then implies that
\begin{align}
  \sum_{\v v} C_{\ell m}\p{\v v} B_{kn}\p{\v v}
  = \delta_{k\ell} \delta_{mn},
  \label{eq:full_recovery}
\end{align}
which is a stronger version of the condition that we found for the tomography problem in Eq.~\eqref{eq:tomo_recovery}.
If, as with the spherical harmonics $Y_{\ell m}$, the functions $B_{\ell m}$ are indexed by an integer degree $\ell\ge0$ that is preserved under rotations of the sphere (which automatically implies integer orders $\abs{m}\le\ell$), and the function $f$ is {\it band-limited} in the sense that $f_{\ell m}=0$ for all $\ell\ge L$, then the sampling problem in Eq.~\eqref{eq:full_recovery} is provably solvable with a suitable choice of $\abs{V}=L^2$ points \cite{freeden2008spherical}.
A solution to the classical signal recovery problem Eq.~\eqref{eq:full_recovery} with $B_{\ell m}=\D^\ell_{m,0}$ automatically solves the quantum tomography problem in Eq.~\eqref{eq:tomo_recovery}, which implies the existence of $d^2$ measurement axes that suffice to reconstruct arbitrary states of $d$-level spin qudit, which are essentially ``band-limited'' at degree $\ell<d$.

Moreover, for any fixed degree $\ell$, the problem in Eq.~\eqref{eq:tomo_recovery} is equivalent to recovery of a function in the degree-$\ell$ subspace of $L^2\p{\Omega_2}$, which is possible with $\abs{V}=2\ell+1$ samples (equal to the dimension of the degree-$\ell$ subspace) \cite{freeden2008spherical}.
In our case, the degree $\ell$ takes a maximal value of $\ell_{\t{max}}=d-1$, so state recovery requires at least as many measurement axes as there are transition operators with degree $\ell_{\t{max}}$, namely $2\ell_{\t{max}}+1=2d-1$.

[\red{Note: finishing comments about why one could expect more than $2d-1$ axes to be necessary, and what conditions are needed for $2d-1$ to be sufficient.}]

[\red{I {\it might} be able to prove that $2d-1$ axes always suffice for full recovery.
  See Appendix \ref{sec:recycle}.}]

\vspace{3cm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Tomography protocol and numerical experiments}

[\red{Note: the text below is copy/pasted from a different set of notes.  I still need to adapt this text for this paper.}]

[\red{Think about maximizing the determinant of the coefficient matrix $C_{m\v v}^\ell$.}]

Question: given a collection of projectors $\Pi\equiv\set{\Pi_j}$ and measurement outcomes $M\equiv\set{M_j}$, with $M_j$ an empirical estimate of $\bk{\Pi_j}_\rho=\tr\p{\rho\Pi_j}$, how do we determine whether the measurement data $M$ is sufficient to reconstruct $\rho$?
Answer: reconstruction of $\rho$ is possible if $\Pi$ spans the entire space $\B_n$ of operators on the Hilbert space of an $n$-level spin.
Given an orthonormal basis $\set{Q_\alpha}$ of self-adjoint operators spanning $\B_n$, we can expand
\begin{align}
  \rho = \sum_\alpha \rho_\alpha Q_\alpha,
  &&
  \rho_\alpha \equiv \bk{Q_\alpha}_\rho.
\end{align}
If $\Pi$ spans $\B_n$, then we can find a set of real numbers $c_{\alpha j}\in\RR$ for which
\begin{align}
  Q_\alpha = \sum_j c_{\alpha j} \Pi_j.
\end{align}
An empirical estimate $\tilde\rho$ of $\rho$ is then
\begin{align}
  \tilde\rho = \sum_\alpha \tilde\rho_\alpha Q_\alpha,
\end{align}
where\footnote{This estimate can be improved by defining $\tilde\rho$ as the least-squares fit to the set of linear equations $\tr\p{\tilde\rho\,\Pi_j} = M_j$, and performing maximum-likelihood corrections to $\tilde\rho$ \cite{smolin2012efficient}.}
\begin{align}
  \tilde\rho_\alpha \equiv \sum_j c_{\alpha j} M_j
  \approx \sum_j c_{\alpha j} \bk{\Pi_j}_\rho
  = \bk{Q_\alpha}_\rho
  = \rho_\alpha.
\end{align}
As it turns out, numerical experiments with $n\le20$ suggest that in order to perfurm full tomography of $\rho$, it suffices to measure all projectors $\Pi_{\v v\mu}$ onto states of definite spin $\mu$ along $2n-1$ different axes $\v v$.
The question remains: which axes should we choose?
As a first pass, we can simply choose a set of $2n-1$ random axes, $V\equiv\set{\v v}$, by uniformly sampling points on the sphere.
These random axes define a set of $\p{2n-1}\times n$ projectors $\Pi_V=\set{\Pi^V_j}$, where for shorthand we use a combined index $j=\p{\v v_j,\mu_j}$ that specifies both a measurement axis $\v v_j$ and an outcome $\mu_j$, and for clarity we index objects by $V$ to emphasize their dependence on the choice of measurement axes.
The projectors $\Pi_V$ can be used to construct $n^2$ linearly independent operators by constructing a matrix $B_V$ with entries
\begin{align}
  \sp{B_V}_{ij} = \tr\p{\Pi^V_i \Pi^V_j}.
\end{align}
Indexing the (normalized) eigenvectors and the corresponding eigenvalues of $B_V$ by $\alpha$, respectively $\v x^V_\alpha$ and $m^V_\alpha$, for all $\alpha$ with eigenvalue $m^V_\alpha\ne0$ we can then construct the normalized and linearly independent operators
\begin{align}
  Q^V_\alpha \equiv \f1{\sqrt{m^V_\alpha}} \sum_j x^V_{\alpha j} \Pi^V_j.
\end{align}
To summarize: choosing a set $V$ of $2n-1$ random axes will determine a set of measurement operators $\Pi_V$ from which we can construct a complete basis $\set{Q^V_\alpha}$ of operators for $\B_n$.
However, some choices of axes will be better than others: a good choice of axes $V$ will maximize the amount of information that the corresponding measurement data $M_V=\set{M^V_j}$ will contain about the expectation values $\bk{Q^V_\alpha}_\rho$ used to reconstruct $\rho$.
The amount of information that $M_V$ contains about $\bk{Q^V_\alpha}_\rho$ is determined, in part, by the magnitude of the eigenvalue $m^V_\alpha$: if $m^V_\alpha$ is small, then $M_V$ will generally contain little information about $\bk{Q^V_\alpha}_\rho$, and vice versa\footnote{Think about the limit $m^V_\alpha\to0$, in which case the fact that $Q^V_\alpha$ is normalized implies that the operator $\sum_j x^V_{\alpha j} \Pi^V_j$ must have a vanishing norm.
  A vanishing norm can only occur due to fine-tuned cancellations that make the expectation value of $\sum_j x^V_{\alpha j} \Pi^V_j$ difficult to estimate from noisy data on $\bk{\Pi^V_j}_\rho$.}.
Rather than choosing $2n-1$ axes randomly, we can therefore choose a set of axes that minimizes some cost function $C\p{m_V}$ favoring large eigenvalues $m_V\equiv\set{m^V_\alpha}$.
A reasonable cost function that treats all eigenvalues on equal footing, for example, is
\begin{align}
  C_{\t{inv}}\p{m_V} \equiv \sum_\alpha \f1{m^V_\alpha},
\end{align}
where the sum is implicitly performed over the $n^2$ largest eigenvalues of $B_V$ (as all other eigenvalues are guaranteed to be zero).
Note that any information about $\rho$, obtained either from prior knowledge or preliminary measurement data, can be used to construct an even better choice of measurement axes, leading to the possibility of tailored or adaptive measurement protocols \cite{pereira2018adaptive} that are more efficient in terms of the number of measurements required to estimate $\rho$ to a fixed precision.
We leave the details of tailored and adaptive measurement protocols to future work.

\bibliography{\jobname.bib}

\onecolumngrid
\appendix

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Rotating transition operators}
\label{sec:rotations}

Defining
\begin{align}
  S_\z \equiv \sum_{\mu=-s}^s \mu \op{\mu},
  &&
  S_\pm \equiv \sum_{\mu=-s}^s
  \sqrt{s\p{s+1}-\mu\p{\mu\pm1}} \op{\mu\pm1}{\mu},
\end{align}
as well as
\begin{align}
  S_\x \equiv \f12\p{S_+ + S_-},
  &&
  S_\y \equiv -\f\i2\p{S_+-S_-},
  &&
  \v S \equiv \p{S_\x,S_\y,S_\z},
\end{align}
the spin vector $\v S$ generates rotations of a spin-$s$ system in 3D space.
Specifically, the operator $e^{-\i\theta\v S\cdot\uv n}$ rotates a spin-$s$ system by an angle $\theta$ about the unit vector $\uv n$.
Furthermore, observing that $S_\z=T_{1,0}$ and $S_\pm\propto T_{1,\pm1}$, we can use the product operator expansion of the transition operators (see Appendix \ref{sec:trans_prod}), the properties of Clebsch-Gordan coefficients, and a computer algebra system to simplify the commutators
\begin{align}
  \sp{S_\z,T_{\ell m}} = m\, T_{\ell m},
  &&
  \sp{S_\pm,T_{\ell m}} = \sqrt{\p{\ell\mp m}\p{\ell\pm m+1}}\, T_{\ell,m\pm1},
\end{align}
which implies that $T_{\ell m}$ is a spherical tensor operator, and in particular that its degree $\ell$ is preserved under rotations generated by $\v S$.
For any triplet of Euler angles $\v\omega=\p{\alpha_{\v\omega},\beta_{\v\omega},\gamma_{\v\omega}}$, we can therefore define the fundamental rotation operator
\begin{align}
  R\p{\v\omega} \equiv e^{-\i\alpha_{\v\omega}S_\z} e^{-\i\beta_{\v\omega}S_\y} e^{-\i\gamma_{\v\omega}S_\z},
\end{align}
as well as the corresponding adjoint rotation operator $\R\p{\v\omega}$ defined by $\R\p{\v\omega}\O = R\p{\v\omega} \O R\p{\v\omega}^\dag$, and expand
\begin{align}
  \R\p{\v\omega} T_{\v\omega\ell m}
  = \sum_{n=-\ell}^\ell \D_{nm}^\ell\p{\v\omega} T_{\ell n}.
\end{align}
The adjoint rotation matrix elements $\D_{mn}^\ell\p{\v\omega}\in\CC$ are
\begin{align}
  \D_{mn}^\ell\p{\v\omega}
  &\equiv \obk{T_{\ell m}|\R\p{\v\omega}|T_{\ell n}} \\
  &= \tr\p{T_{\ell m}^\dag e^{-\i\alpha_{\v\omega}S_\z} e^{-\i\beta_{\v\omega}S_\y} e^{-\i\gamma_{\v\omega}S_\z} T_{\ell n}
    e^{\i\gamma_{\v\omega}S_\z} e^{\i\beta_{\v\omega}S_\y} e^{\i\alpha_{\v\omega}S_\z}} \\
  &= e^{-\i\alpha_{\v\omega} m} \,\dd_{mn}^\ell\p{\beta_{\v\omega}}\, e^{-\i\gamma_{\v\omega} n},
\end{align}
and the {\it reduced} adjoint rotation matrix elements $\dd_{mn}^\ell\p{\beta}\in\RR$ are
\begin{align}
  \dd_{mn}^\ell\p{\beta}
  &\equiv \D_{mn}^\ell\p{0,\beta,0} \\
  &= \tr\p{T_{\ell m}^\dag e^{-\i\beta S_\y} T_{\ell n} e^{\i\beta S_\y}} \\
  &= \f{2\ell+1}{2s+1} \sum_{\mu,\nu=-s}^s
  \bk{s\mu;\ell m|s,\mu+m} \bk{s\nu;\ell n|s,\nu+n} \times
  \tr\p{\op{\mu}{\mu+m} e^{-\i\beta S_\y}
    \op{\nu+n}{\nu} e^{\i\beta S_\y}} \\
  &= \f{2\ell+1}{2s+1} \sum_{\mu,\nu=-s}^s
  \bk{s\mu;\ell m|s,\mu+m} \bk{s\nu;\ell n|s,\nu+n} \times
  d_{\mu\nu}^s\p{\beta} d_{\mu+m,\nu+n}^s\p{\beta}.
\end{align}
Here $d_{\mu\nu}^s\p{\beta}\equiv D^s_{\mu\nu}\p{0,\beta,0}\in\RR$ are
matrix elements of a reduced fundamental (Wigner) rotation matrix.
Note that the above product of reduced fundamental rotation matrix elements can be expanded as \cite{rose1957elementary}
\begin{align}
  d_{\mu\nu}^s\p{\beta} d_{\mu+m,\nu+n}^s\p{\beta}
  = \sum_{L=0}^{2s} \bk{s\mu;s,\mu+m|L,2\mu+m}
  \bk{s\nu;s,\nu+n|L,2\nu+n} d_{2\mu+m,2\nu+n}^L\p{\beta}.
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Rotation matrix elements as orthogonal functions}
\label{sec:ortho}

As functions of $\v\omega$, the fundamental rotation matrix elements $D_{mn}^\ell$ and satisfy the orthogonality relations \cite{brown2003rotational}
\begin{align}
  \int_{\Omega_3} \d\Omega_3\p{\v\omega}
  D^{\ell}_{mn}\p{\v\omega}^* D^{\ell'}_{m'n'}\p{\v\omega}
  = \f{8\pi^2}{2\ell+1} \delta_{\ell\ell'} \delta_{mm'} \delta_{nn'},
  \label{eq:orthogonal}
\end{align}
where integration is performed over all $\v\omega\in\Omega_3\simeq[0,2\pi]\times[0,\pi]\times[0,2\pi]$ with integration measure $\d\Omega_3\p{\v\omega}\equiv\d\alpha_{\v\omega} \d\beta_{\v\omega} \d\gamma_{\v\omega} \sin\beta_{\v\omega}$.
This orthogonality relation is a consequence of the fact that, for a fixed degree $\ell$, the matrix
\begin{align}
  D^\ell \equiv \sum_{m,n} D^\ell_{mn} \op{m}{n}
\end{align}
is an irreducible unitary matrix representation of SO(3), where by unitary we mean that
\begin{align}
  \sum_k D^\ell_{km}\p{\v\omega}^* D^\ell_{kn}\p{\v\omega} = \delta_{mn}
  ~\t{for all}~\v\omega\in\Omega_3\simeq\SO(3).
\end{align}
Unitarity of $D^\ell$ follows from the fact that the matrix $D^\ell\p{\v\omega}$ is essentially the degree-$\ell$ block of the block-diagonal fundamental rotation matrix $R\p{\v\omega}$, which is itself unitary.
The Schur orthogonality relations for irreducible unitary representations of compact Lie groups then imply that \cite{cornwell1997group}
\begin{align}
  \int_{\Omega_3} \d\Omega_3\p{\v\omega}
  D^{\ell}_{mn}\p{\v\omega}^* D^{\ell'}_{m'n'}\p{\v\omega}
  = \f{\abs{\Omega_3}}{\dim\p{D^\ell}}
  \delta_{\ell\ell'} \delta_{mm'} \delta_{nn'},
\end{align}
where $\dim\p{D^\ell}=2\ell+1$ is the dimension of $D^\ell$, and analogously
\begin{align}
  \abs{\Omega_3} \equiv \int_{\Omega_3} d\Omega_3\p{\v\omega} = 8\pi^2.
\end{align}
Orthogonality relations for adjoint matrix elements $\D^\ell_{mn}$ are derived identically from the fact that the matrix
\begin{align}
  \D^\ell \equiv \sum_{m,n} \D^\ell_{mn} \op{m}{n}
\end{align}
is an irreducible unitary representation of SO(3).
This property of $\D^\ell$ follows from the fact that it is an adjoint representation of SO(3), which is isomorphic to the standard representation $D^\ell$ \cite{hall2015lie}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Recycling sampling points on the sphere}
\label{sec:recycle}

[\red{Note: I have not finished writing up the proof below.
  I was previously sure that it works, but I ran into an error that I am trying to fix, so I am not so sure at the moment.}]

Consider a set of points $V=\set{\v v}$ on the sphere, and define $L^2_\ell\p{\Omega}$ as the subspace of square-integrable functions on the sphere that can be written as a sum of degree-$\ell$ spherical harmonics, which is to say that a function $f_\ell\in L^2_\ell\p{\Omega}$ admits the decomposition
\begin{align}
  f_\ell = \sum_{m=-\ell}^\ell f_{\ell m} Y_{\ell m},
\end{align}
where $f_{\ell m}\in\CC$ are scalars and $Y_{\ell m}$ is the spherical harmonic of degree $\ell$ and order $m$.
Here, we prove the following:
\begin{theorem}
  If $V$ enables the recovery of arbitrary functions $f_\ell\in L^2_\ell\p{\Omega}$, which is to say that $f_\ell$ is uniquely determined by its values $f_\ell\p{\v v}$ at all $\v v\in V$, then $V$ also enables the recovery of arbitrary functions $f_k\in L^2_k\p{\Omega}$ with $k<\ell$.
  \label{thm:recycling}
\end{theorem}
Consider the equivalent, contrapositive statement: if a set of points $V$ is insufficient to recover functions $f_k\in L^2_k\p{\Omega}$, then $V$ is insufficient to recover functions $f_\ell\in L^2_\ell\p{\Omega}$ with $\ell>k$.
If we can prove this statement for $k=\ell-1$, then we essentially prove it for all $k<\ell$ by induction, which in turn proves theorem \ref{thm:recycling}.

We first clarify necessary and sufficient conditions for $V$ to enable the recovery of arbitrary functions $f_k\in L^2_k\p{\Omega}$.
Recovery requires the existence of coefficients $C_{km}\p{\v v}$ for which
\begin{align}
  f_{km} = \sum_{\v v\in V} C_{km} f_k\p{v}
  = \sum_{\v v,n} C_{km}\p{\v v} f_{kn} Y_{kn}\p{\v v},
\end{align}
which implies that
\begin{align}
  \sum_{\v v\in V} C_{km}\p{\v v} Y_{kn}\p{\v v} = \delta_{mn}.
  \label{eq:degree_recovery}
\end{align}
Coefficients $C_{km}\p{\v v}$ satisfying Eq.~\eqref{eq:degree_recovery} are essentially a left-inverse of the $\abs{V}\times\p{2k+1}$ matrix
\begin{align}
  \Y_k\p{V} \equiv
  \begin{pmatrix}
    Y_{k,-k}\p{\v v_1} & \cdots & Y_{kk}\p{\v v_1} \\
    \vdots & \ddots & \vdots \\
    Y_{k,-k}\p{\v v_{\abs{V}}} & \cdots & Y_{kk}\p{\v v_{\abs{V}}}
  \end{pmatrix},
\end{align}
which exists if and only if $\Y_k\p{V}$ has maximal rank, $2k+1$.
Note that the existence of such an inverse is entirely a property of $V$, independent of the choice of basis for $L^2_k\p{\Omega}$ \cite{freeden2008spherical}, which is why we are free to consider the spherical harmonics $Y_{km}$ here in place of the adjoint rotation matrix elements $\D^k_{m,0}$ that appear in the main text.

We now assume that $V$ is insufficient to recover functions in $L^2_{\ell-1}\p{\Omega}$.
In this case, the matrix $\Y_{\ell-1}\p{V}$ does not have maximal rank, which implies that its columns are linearly dependent.
Linear dependence of the columns of $\Y_{\ell-1}\p{V}$ in turn implies the existence of non-trivial (i.e.~not-all-zero) coefficients $c_{\ell-1,m}$ for which
\begin{align}
  \sum_{m=-\ell+1}^{\ell-1} c_{\ell-1,m} Y_{\ell-1,m}\p{\v v}
  = 0 ~\t{for all}~ \v v\in V.
\end{align}
Using the fact that $Y_{km}\p{\v v}\propto e^{\i m\alpha_{\v v}} P_{km}\p{\cos\beta_{\v v}}$, where $\alpha_{\v v},\beta_{\v v}$ are respectively the azimuthal and polar angles of $\v v$ and $P_{km}$ is an associated Legendre polynomial, we can say that
\begin{align}
  \sum_{m=-\ell+1}^{\ell-1} \tilde c_{\ell-1,m} e^{\i m\alpha_{\v v}} P_{\ell-1,m}\p{\cos\beta_{\v v}}
  = 0 ~\t{for all}~ \v v\in V,
  &&
  \tilde c_{\ell m} \equiv c_{\ell m}
  \times \sqrt{\f{4\pi\p{\ell+m}!}{\p{2\ell+1}\p{\ell-m}!}}.
\end{align}
Assuming without loss of generality that $\cos\beta_{\v v}\ne1$ for all $\v v\in V$, we can use recurrence relations of the associated Legendre polynomials to say that
\begin{align}
  \sum_{m=-\ell+1}^{\ell-1} \sum_{\sigma\in\set{\pm1}}
  e^{\i m\alpha_{\v v}} \tilde d_{\ell m\sigma}
  P_{\ell m+\sigma}\p{\cos\beta_{\v v}}
  = 0 ~\t{for all}~ \v v\in V,
\end{align}
where
\begin{align}
  \tilde d_{\ell m,+} \equiv -\f{\tilde c_{\ell-1,m}}{2m},
  &&
  \tilde d_{\ell m,-} \equiv -\f{\tilde c_{\ell-1,m}}{2m}
  \p{\ell-m+1} \p{\ell-m+2}.
\end{align}
[\red{what if $m=0$?}] Substituting back the spherical harmonics in place of the associated Legendre polynomials, we thus find that
\begin{align}
  \sum_{m=-\ell}^\ell \sum_{\sigma\in\set{\pm 1}}
  d_{\ell m\sigma} e^{-\i\sigma\alpha_{\v v}} Y_{\ell m}\p{\v v} = 0
  ~\t{for all}~ \v v\in V,
  &&
  d_{\ell m\sigma} \equiv \tilde d_{\ell,m-\sigma,\sigma}
  \times \sqrt{\f{\p{2\ell+1}\p{\ell-m}!}{4\pi\p{\ell+m}!}},
\end{align}
where we define $\tilde d_{\ell\ell\sigma} = \tilde d_{\ell,-\ell,\sigma} = 0$.

\vspace{.5cm}

[\red{proof incomplete}]

\vspace{3cm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Transition operator product expansion}
\label{sec:trans_prod}

The transition operators on the $d$-dimensional Hilbert space of a spin-$s$ system (with $s\equiv\frac{d-1}{2}$) are defined by
\begin{align}
  T_{\ell m} = \sqrt{\f{2\ell+1}{2s+1}} \sum_{\mu,\nu=-s}^s
  \bk{s\mu;\ell m|s\nu} \op{\nu}{\mu},
\end{align}
where $\bk{s\mu;\ell m|s\nu}$ is a Clebsh-Gordan coefficient that enforces $\ell\in\set{0,1,\cdots,2s}$ and $m\in\set{-\ell,-\ell+1,\cdots,\ell}$.
We wish to compute the coefficients of the operator product expansion
\begin{align}
  T_{\ell_1 m_1} T_{\ell_2 m_2}
  = \sum_{L,M} f_{\ell_1 m_1;\ell_2 m_2}^{LM} T_{LM},
  &&
  f_{\ell_1 m_1;\ell_2 m_2}^{LM}
  \equiv \obk{T_{LM} | T_{\ell_1 m_1} T_{\ell_2 m_2}}.
\end{align}
Using the symmetry properties of Clebsch-Gordan coefficients, namely
\begin{align}
  \bk{\ell_1 m_1; \ell_2 m_2| \ell_3 m_3}
  &= \p{-1}^{\ell_2+m_2} \sqrt{\f{2\ell_3+1}{2\ell_1+1}}
  \bk{\ell_3,-m_3; \ell_2 m_2| \ell_1,-m_1} \\
  \bk{\ell_1 m_1; \ell_2 m_2| \ell_3 m_3}
  &= \p{-1}^{\ell_1+\ell_2-\ell_3}
  \bk{\ell_1,-m_1; \ell_2,-m_2| \ell_3,-m_3},
\end{align}
we can find that the transition operators transform under conjugation as
\begin{align}
  T_{\ell m}^\dag
  = \sqrt{\f{2\ell+1}{2s+1}}
  \sum_{\mu,\nu} \p{-1}^m \bk{s\nu;\ell,-m|s\mu} \op{\mu}{\nu}
  = \p{-1}^m T_{\ell,-m},
\end{align}
which implies that
\begin{align}
  f_{\ell_1 m_1;\ell_2 m_2}^{LM}
  = \p{-1}^M \sqrt{\f{\p{2L+1}\p{2\ell_1+1}\p{2\ell_2+1}}
    {\p{2s+1}\p{2s+1}\p{2s+1}}}
  \sum_{\mu,\nu,\rho} \bk{s\nu;L,-M|s\mu}
  \bk{s\rho;\ell_1m_1|s\nu} \bk{s\mu;\ell_2m_2|s\rho}.
\end{align}
Replacing the Clebsch-Gordan coefficients by the Wigner 3-$j$ symbols with the identity
\begin{align}
  \bk{\ell_1 m_1; \ell_2 m_2| \ell_3 m_3}
  = \p{-1}^{2\ell_2} \p{-1}^{\ell_3-m_3} \sqrt{\p{2\ell_3+1}}
  \begin{pmatrix}
    \ell_3 & \ell_2 & \ell_1 \\
    -m_3 & m_2 & m_1
  \end{pmatrix},
\end{align}
we can use the fact that $2\ell_2$ is (in our case) always even to expand
\begin{align}
  f_{\ell_1 m_1;\ell_2 m_2}^{LM}
  = \p{-1}^M \sqrt{\p{2L+1}\p{2\ell_1+1}\p{2\ell_2+1}}
  \sum_{\mu,\nu,\rho} \p{-1}^{3s-\mu-\nu-\rho}
  \begin{pmatrix}
    s & L & s \\
    -\mu & -M & \nu
  \end{pmatrix}
  \begin{pmatrix}
    s & \ell_1 & s \\
    -\nu & m_1 & \rho
  \end{pmatrix}
  \begin{pmatrix}
    s & \ell_2 & s \\
    -\rho & m_2 & \mu
  \end{pmatrix}.
\end{align}
This sum can be simplified by the introduction of Wigner 6-$j$ symbols, giving us
\begin{align}
  f_{\ell_1 m_1;\ell_2 m_2}^{LM}
  &= \p{-1}^{2s+M} \sqrt{\p{2L+1}\p{2\ell_1+1}\p{2\ell_2+1}}
  \begin{pmatrix}
    L & \ell_1 & \ell_2 \\
    M & -m_1 & -m_2
  \end{pmatrix}
  \begin{Bmatrix}
    L & \ell_1 & \ell_2 \\
    s & s & s
  \end{Bmatrix} \\
  &= \p{-1}^{2s+L} \sqrt{\p{2\ell_1+1}\p{2\ell_2+1}}
  \bk{\ell_1 m_1; \ell_2 m_2| LM}
  \begin{Bmatrix}
    \ell_1 & \ell_2 & L \\
    s & s & s
  \end{Bmatrix}.
\end{align}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2020-10-14 sun_phases/spin_qudit_tomo.tex
remove failed attempt to prove that only 2d-1 axes are always sufficient for tomography

@@
 
-[\red{I {\it might} be able to prove that $2d-1$ axes always suffice for full recovery.
-  See Appendix \ref{sec:recycle}.}]
-
 \vspace{3cm}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Recycling sampling points on the sphere}
-\label{sec:recycle}
-
-[\red{Note: I have not finished writing up the proof below.
-  I was previously sure that it works, but I ran into an error that I am trying to fix, so I am not so sure at the moment.}]
-
-Consider a set of points $V=\set{\v v}$ on the sphere, and define $L^2_\ell\p{\Omega}$ as the subspace of square-integrable functions on the sphere that can be written as a sum of degree-$\ell$ spherical harmonics, which is to say that a function $f_\ell\in L^2_\ell\p{\Omega}$ admits the decomposition
-\begin{align}
-  f_\ell = \sum_{m=-\ell}^\ell f_{\ell m} Y_{\ell m},
-\end{align}
-where $f_{\ell m}\in\CC$ are scalars and $Y_{\ell m}$ is the spherical harmonic of degree $\ell$ and order $m$.
-Here, we prove the following:
-\begin{theorem}
-  If $V$ enables the recovery of arbitrary functions $f_\ell\in L^2_\ell\p{\Omega}$, which is to say that $f_\ell$ is uniquely determined by its values $f_\ell\p{\v v}$ at all $\v v\in V$, then $V$ also enables the recovery of arbitrary functions $f_k\in L^2_k\p{\Omega}$ with $k<\ell$.
-  \label{thm:recycling}
-\end{theorem}
-Consider the equivalent, contrapositive statement: if a set of points $V$ is insufficient to recover functions $f_k\in L^2_k\p{\Omega}$, then $V$ is insufficient to recover functions $f_\ell\in L^2_\ell\p{\Omega}$ with $\ell>k$.
-If we can prove this statement for $k=\ell-1$, then we essentially prove it for all $k<\ell$ by induction, which in turn proves theorem \ref{thm:recycling}.
-
-We first clarify necessary and sufficient conditions for $V$ to enable the recovery of arbitrary functions $f_k\in L^2_k\p{\Omega}$.
-Recovery requires the existence of coefficients $C_{km}\p{\v v}$ for which
-\begin{align}
-  f_{km} = \sum_{\v v\in V} C_{km} f_k\p{v}
-  = \sum_{\v v,n} C_{km}\p{\v v} f_{kn} Y_{kn}\p{\v v},
-\end{align}
-which implies that
-\begin{align}
-  \sum_{\v v\in V} C_{km}\p{\v v} Y_{kn}\p{\v v} = \delta_{mn}.
-  \label{eq:degree_recovery}
-\end{align}
-Coefficients $C_{km}\p{\v v}$ satisfying Eq.~\eqref{eq:degree_recovery} are essentially a left-inverse of the $\abs{V}\times\p{2k+1}$ matrix
-\begin{align}
-  \Y_k\p{V} \equiv
-  \begin{pmatrix}
-    Y_{k,-k}\p{\v v_1} & \cdots & Y_{kk}\p{\v v_1} \\
-    \vdots & \ddots & \vdots \\
-    Y_{k,-k}\p{\v v_{\abs{V}}} & \cdots & Y_{kk}\p{\v v_{\abs{V}}}
-  \end{pmatrix},
-\end{align}
-which exists if and only if $\Y_k\p{V}$ has maximal rank, $2k+1$.
-Note that the existence of such an inverse is entirely a property of $V$, independent of the choice of basis for $L^2_k\p{\Omega}$ \cite{freeden2008spherical}, which is why we are free to consider the spherical harmonics $Y_{km}$ here in place of the adjoint rotation matrix elements $\D^k_{m,0}$ that appear in the main text.
-
-We now assume that $V$ is insufficient to recover functions in $L^2_{\ell-1}\p{\Omega}$.
-In this case, the matrix $\Y_{\ell-1}\p{V}$ does not have maximal rank, which implies that its columns are linearly dependent.
-Linear dependence of the columns of $\Y_{\ell-1}\p{V}$ in turn implies the existence of non-trivial (i.e.~not-all-zero) coefficients $c_{\ell-1,m}$ for which
-\begin{align}
-  \sum_{m=-\ell+1}^{\ell-1} c_{\ell-1,m} Y_{\ell-1,m}\p{\v v}
-  = 0 ~\t{for all}~ \v v\in V.
-\end{align}
-Using the fact that $Y_{km}\p{\v v}\propto e^{\i m\alpha_{\v v}} P_{km}\p{\cos\beta_{\v v}}$, where $\alpha_{\v v},\beta_{\v v}$ are respectively the azimuthal and polar angles of $\v v$ and $P_{km}$ is an associated Legendre polynomial, we can say that
-\begin{align}
-  \sum_{m=-\ell+1}^{\ell-1} \tilde c_{\ell-1,m} e^{\i m\alpha_{\v v}} P_{\ell-1,m}\p{\cos\beta_{\v v}}
-  = 0 ~\t{for all}~ \v v\in V,
-  &&
-  \tilde c_{\ell m} \equiv c_{\ell m}
-  \times \sqrt{\f{4\pi\p{\ell+m}!}{\p{2\ell+1}\p{\ell-m}!}}.
-\end{align}
-Assuming without loss of generality that $\cos\beta_{\v v}\ne1$ for all $\v v\in V$, we can use recurrence relations of the associated Legendre polynomials to say that
-\begin{align}
-  \sum_{m=-\ell+1}^{\ell-1} \sum_{\sigma\in\set{\pm1}}
-  e^{\i m\alpha_{\v v}} \tilde d_{\ell m\sigma}
-  P_{\ell m+\sigma}\p{\cos\beta_{\v v}}
-  = 0 ~\t{for all}~ \v v\in V,
-\end{align}
-where
-\begin{align}
-  \tilde d_{\ell m,+} \equiv -\f{\tilde c_{\ell-1,m}}{2m},
-  &&
-  \tilde d_{\ell m,-} \equiv -\f{\tilde c_{\ell-1,m}}{2m}
-  \p{\ell-m+1} \p{\ell-m+2}.
-\end{align}
-[\red{what if $m=0$?}] Substituting back the spherical harmonics in place of the associated Legendre polynomials, we thus find that
-\begin{align}
-  \sum_{m=-\ell}^\ell \sum_{\sigma\in\set{\pm 1}}
-  d_{\ell m\sigma} e^{-\i\sigma\alpha_{\v v}} Y_{\ell m}\p{\v v} = 0
-  ~\t{for all}~ \v v\in V,
-  &&
-  d_{\ell m\sigma} \equiv \tilde d_{\ell,m-\sigma,\sigma}
-  \times \sqrt{\f{\p{2\ell+1}\p{\ell-m}!}{4\pi\p{\ell+m}!}},
-\end{align}
-where we define $\tilde d_{\ell\ell\sigma} = \tilde d_{\ell,-\ell,\sigma} = 0$.
-
-\vspace{.5cm}
-
-[\red{proof incomplete}]
-
-\vspace{3cm}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Transition operator product expansion}

### 2020-10-20 sun_phases/sun_dynamics_notes.tex
changed some conventions, and re-ordered table listing possible drive hamiltonians

@@
 Perhaps more interestingly, we can in principle prepare inhomogeneous product states in which the state of each spin is correlated with its corresponding single-body term of the spin Hamiltonian in \eqref{eq:spin_pair}.
-We are currently working on characterizing the sort of inhomogeneous states that we can prepare.
+We are currently working on characterizing the sort of inhomogeneous states that we can prepare (see Appendix \ref{sec:init_states}).
 
@@
 
-% Motivations / background:
-% https://arxiv.org/pdf/1901.06282.pdf
-% https://arxiv.org/abs/1911.00099
-% https://arxiv.org/abs/2005.10910
-% https://journals.aps.org/pra/abstract/10.1103/PhysRevA.86.012118
-% https://journals.aps.org/pra/abstract/10.1103/PhysRevA.94.010301
-% https://arxiv.org/pdf/2003.08668.pdf
-% https://www.osapublishing.org/ol/abstract.cfm?uri=ol-44-10-2558
-
 Here we discuss the task of performing full tomography on the average reduced single-particle density operator of a collection of nuclear spins.
@@
 To write out $s_{\v v\pm}$ in full, we first identify the operators $s_\x,s_\y,s_\z$ that generate rotations of an $n$-level spin in 3D space, and define $s_\pm\equiv s_\x\pm\i s_\y$.
-Denoting the polar and azimuthal angles of $\v v$ respectively by $\alpha_{\v v}$ and $\beta_{\v v}$, we can use the fact that $s_\x,s_\y,s_\z$ form an $\su(2)$ algebra to expand
+Denoting the azimuthal and polar angles of $\v v$ respectively by $\alpha_{\v v}$ and $\beta_{\v v}$, we can use the fact that $s_\x,s_\y,s_\z$ form an $\su(2)$ algebra to expand
 \begin{align}
   s_{\v v\pm}
-  &= e^{-\i\beta_{\v v}s_\z} e^{-\i\alpha_{\v v}s_\y}
-  s_\pm e^{\i\alpha_{\v v}s_\y} e^{\i\beta_{\v v}s_\z} \\
-  &= -\sin\alpha_{\v v}\, s_\z
-  + \f12 \p{\cos\alpha_{\v v}+1} e^{\mp\i\beta_{\v v}} s_\pm
-  + \f12 \p{\cos\alpha_{\v v}-1} e^{\pm\i\beta_{\v v}} s_\mp.
+  &= e^{-\i\alpha_{\v v}s_\z} e^{-\i\beta_{\v v}s_\y}
+  s_\pm e^{\i\beta_{\v v}s_\y} e^{\i\alpha_{\v v}s_\z} \\
+  &= -\sin\beta_{\v v}\, s_\z
+  + \f12 \p{\cos\beta_{\v v}+1} e^{\mp\i\alpha_{\v v}} s_\pm
+  + \f12 \p{\cos\beta_{\v v}-1} e^{\pm\i\alpha_{\v v}} s_\mp.
 \end{align}
@@
   - \p{\Omega_+ + \Omega_-}^2 s_{\tilde\eta_0,\x}^2
-  - \p{\Omega_+ - \Omega_-}^2 s_{\tilde\eta_0,\y}^2
-  ,
+  - \p{\Omega_+ - \Omega_-}^2 s_{\tilde\eta_0,\y}^2,
   \label{eq:drive_single}
@@
   \caption{Drive Hamiltonians that can be implemented with various amplitude-matching conditions when all $\eta_m=0$.
-    Here $\v s_{\t{3D}}^2=s_\x^2+s_\y^2+s_\z^2=I\p{I+1}$ is a constant.}
+    Here $\v s_{\t{3D}}^2=s_\x^2+s_\y^2+s_\z^2=I\p{I+1}$ is a constant and $\sigma\in\set{+1,-1}$.
+For each $m\in\set{0,+1,-1}$, a drive with $\abs*{\tilde\Omega_m}=1$ and $\tilde\Omega_n=0$ for $n\ne m$ commutes with drives in which $\tilde\Omega_m=0$ and $\abs*{\tilde\Omega_n}=1$ for $n\ne m$.}
   \begin{tabular}{c|c}
@@
     \\ \hline
-    $\p{0,1,\pm1}$ & $\pm s_\z + s_\z^2 - \v s_{\t{3D}}^2$
+    $\p{0,1,\sigma}$ & $\sigma s_\z + s_\z^2 - \v s_{\t{3D}}^2$
     \\ \hline
-    $\p{1,0,\pm1}$ & $\pm s_\x + s_\x^2 - \v s_{\t{3D}}^2$
+    $\p{1,0,\sigma}$ & $\sigma s_\x + s_\x^2 - \v s_{\t{3D}}^2$
     \\ \hline
-    $\p{1,\pm1,0}$
-    & $\pm\p{s_\z s_\x+s_\x s_\z} + s_\y^2 - \v s_{\t{3D}}^2$
+    $\p{1,\sigma,0}$
+    & $\sigma\p{s_\z s_\x+s_\x s_\z} + s_\y^2 - \v s_{\t{3D}}^2$
+    \\ \hline
+    $\p{1,\sigma,\pm\sigma}$
+    & $\pm s_\z \pm \sigma s_\x
+    + \sigma \p{s_\z s_\x + s_\x s_\z} - \v s_{\t{3D}}^2$
   \end{tabular}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Initial states}
+\label{sec:init_states}
+
+Here we explore the preparation of initial states that are correlated with the single-particle terms of the spin Hamiltonian in Eqs.~\eqref{eq:spin_cos} and \eqref{eq:spin_pair}, namely
+\begin{align}
+  H_{\t{SOC}} \equiv -2 J \sum_{q,\mu} \cos\p{q+\mu\phi} s_{\mu\mu q}.
+\end{align}
+
 \bibliography{sun_dynamics_notes.bib}

### 2020-10-20 sun_phases/spin_qudit_tomo.tex
added text to the intro added explicit mapping between existing solutions to the signal recovery problem and solutions to the tomography problem

@@
 \renewcommand{\sp}[1]{\left[#1\right]} % square parenthesis
-\renewcommand{\set}[1]{\left\{#1\right\}} % curly parenthesis
+\renewcommand{\set}[1]{\{#1\}} % curly parenthesis
+\renewcommand{\Set}[1]{\left\{#1\right\}} % curly parenthesis
 \newcommand{\bk}{\Braket} % shorthand for braket notation
@@
   We consider the task of performing quantum state tomography on a $d$-state spin qudit, using only measurements of spin projection onto different quantization axes.
-  By mapping this quantum tomography problem onto the well-studied classical problem of signal recovery on the sphere, we prove that full reconstruction of arbitrary mixed states is possible, and requires a minimum number of measurement axes, $r_d$, that is bounded by $2d-1\le r_d\le d^2$.
+  By mapping this quantum tomography problem onto the classical problem of signal recovery on the sphere, we prove that full reconstruction of arbitrary mixed states is possible, and requires a minimum number of measurement axes, $r_d$, that is bounded by $2d-1\le r_d\le d^2$.
   We then provide an explicit optimized tomographic protocol, as well as numerical experiments to show that $r_d=2d-1$ for $d\le20$.
@@
 
-Quantum state tomography, or the task of reconstructing a quantum state by collecting and processing measurement data, is an essential algorithmic primitive for quantum sensing, quantum simulation, and quantum information processing.
+Quantum state tomography, or the task of reconstructing a quantum state by collecting and processing measurement data, is an essential primitive for quantum sensing, quantum simulation, and quantum information processing.
 The central importance of quantum state tomography has led to the development of a variety of techniques based on compressed sensing \cite{gross2010quantum}, matrix product states \cite{cramer2010efficient}, % confidence regions \cite{christandl2012reliable},
 maximum-likelihood estimation \cite{smolin2012efficient}, % linear \cite{qi2013quantum} and least-squares \cite{opatrny1997leastsquares} regression,
-Bayesian inference \cite{huszar2012adaptive}, and neural networks \cite{torlai2018neuralnetwork}, among others that are too numerous to list here.
+Bayesian inference \cite{huszar2012adaptive}, and neural networks \cite{torlai2018neuralnetwork}, among others.
 These techniques are typically developed in a general, information-theoretic setting, and make minimal assumptions about the physical medium of a quantum state.
@@
 
-[\red{Discuss spin qudits as quantum resources for quantum sensing \cite{stefano2019set}, quantum simulation \cite{banerjee2013atomic, cazalilla2014ultracold, rico2018nuclear}, and quantum information processing \cite{albert2020robust, gross2020encoding}}]
+Due in part to advancements in experimental capabilities to address nuclear spins in ultracold atomic systems [\red{CITE}], as well as developments in molecular cooling techniques [\red{CITE}], a particular setting of growing interest is the spin qudit, which describes a quantum angular momentum degree of freedom.
+Multilevel spin qudits can provide advantages for quantum sensing \cite{evrard2019enhanced}, enable quantum simulations of SU($N$) magnetism \cite{banerjee2013atomic, cazalilla2014ultracold}, and offer unique capabilities for quantum error correction \cite{albert2020robust, gross2020encoding}.
+In all cases, quantum state tomography is necessary to take full advantage of a spin qudit.
 
-[\red{Discuss previous work on qudit tomography \cite{thew2002qudit, flammia2005minimal, salazar2012quantum, sosa-martinez2017quantum, ha2018minimal, evrard2019enhanced, stefano2019set, palici2020oam}}]
+The problem of qudit tomography in particular is not new, with an extensive literature on a variety of techniques \cite{manko1997spin, thew2002qudit, flammia2005minimal, salazar2012quantum, sosa-martinez2017quantum, ha2018minimal, evrard2019enhanced, stefano2019set, palici2020oam}.
+However, existing proposals either require the capability to perform essentially arbitrary operations on a qudit \cite{thew2002qudit, flammia2005minimal, salazar2012quantum, sosa-martinez2017quantum, ha2018minimal, stefano2019set, palici2020oam}, or rely on infinite-dimensional representations of a quantum spin \cite{manko1997spin, evrard2019enhanced}, resulting in tomographic protocols that can be highly inefficient or even impossible in practice.
+In particular, the protocols based on infinite-dimensional representations of a quantum spin require only the capability to measure spin projection onto different spatial axes, which is usually straightforward with a spin qudit.
+Nonetheless, these protocols obfuscate the requirements for performing full state tomography, provide no straightforward guarantees of state reconstruction fidelity, and extract only a vanishing faction of the information contained in measurement data.
 
-[\red{Discuss practical limitations with spin qudits: the operations necessary for existing tomographic protocols may be inaccessible, difficult to implement, or otherwise more costly than simple spin projection measurements.}]
+In this work, we provide a new and simple framework for performing spin qudit tomography.
+We begin by mapping this quantum tomography problem onto the well-studied classical problem of signal recovery on the sphere, thereby importing a host of existing literature and machinery that has been developed for this task \cite{mcewen2011novel, mcewen2011sampling, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
 
-[\red{Mention previous work on spin tomography \cite{manko1997spin}.
-Discuss connection to quantum state tomography through the discrete Wigner function \cite{leonhardt1995quantumstate, leonhardt1996discrete}.}]
-
-[\red{Discuss connection to signal recovery on the sphere.
-Original $\sim4d^2$ signaling theorem \cite{driscoll1994computing}; its $\sim2d^2$ refinement \cite{mcewen2011novel}; a nice review of these methods \cite{mcewen2011sampling}; an empirically accurate (albeit without rigorous proofs/guarantees) $d^2$ sampling method \cite{khalid2014optimaldimensionality}; current status of the theory behind sparse signal recovery with random sampling \cite{rauhut2011sparse} (note: existing bounds are {\it not} believed to be tight); nice numerical method for sparse signal recovery \cite{alem2012sparse}; useful textbook references \cite{freeden2008spherical, freeden2018spherical}.}]
+[\red{finish intro, provide outline of the paper}]
 
@@
 \end{align}
-The adjoint rotation matrix elements $\D_{mn}^\ell\p{\v\omega}\equiv\obk{T_{\ell m}|\R\p{\v\omega}|T_{\ell n}}$ are an analog of the fundamental (Wigner) rotation matrix elements $D_{mn}^\ell\p{\v\omega}\equiv\bk{\ell m|R\p{\v\omega}|\ell n}$ that govern rotations of spin-$\ell$ particles, where $\ket{\ell m}$ denotes the state of a spin-$\ell$ qudit with spin projection $m$ onto a quantization axis.
+The adjoint rotation matrix elements $\D_{mn}^\ell\p{\v\omega}\equiv\obk{T_{\ell m}|\R\p{\v\omega}|T_{\ell n}}$ are an analog of the fundamental (Wigner) rotation matrix elements $D_{mn}^\ell\p{\v\omega}\equiv\bk{\ell m|R\p{\v\omega}|\ell n}$ that govern rotations of spin-$\ell$ particles and spherical harmonics $Y_{\ell m}$, where $\ket{\ell m}$ denotes the state of a spin-$\ell$ qudit with spin projection $m$ onto a quantization axis.
 Throughout this work, we will primarily consider rotations determined by points on the 2-sphere $\Omega_2$ at $\gamma_{\v\omega}=0$.
@@
   = \f{8\pi^2}{2\ell+1} \delta_{\ell\ell'} \delta_{mm'} \delta_{nn'},
-  \label{eq:ortho}
+  \label{eq:ortho_full}
 \end{align}
 where integration is performed over all $\v\omega\in\Omega_3\simeq[0,2\pi]\times[0,\pi]\times[0,2\pi]$ with integration measure $\d\Omega_3\p{\v\omega} \equiv \d\alpha_{\v\omega} \d\beta_{\v\omega} \d\gamma_{\v\omega} \sin\beta_{\v\omega}$.
-In particular, the restriction of $D^\ell_{m,0}$ to the 2-sphere $\Omega_2$ at $\gamma_{\v\omega}=0$ essentially recovers (up to normalization factors and complex conjugation) the spherical harmonics $Y_{\ell m}$.
-The same orthogonality relations in Eq.~\eqref{eq:ortho} are inherited by the rotation matrix elements $\D_{mn}^\ell$ through the fact $\R\p{\v\omega}$ is the adjoint action of $R\p{\v\omega}$ (see Appendix \ref{sec:ortho}), and the functions $\set{\D_{m,0}^\ell}$ restricted to $\Omega_2$ are similarly an orthogonal basis for the space $L^2\p{\Omega_2}$ of square-integrable functions on the sphere.
+In particular, the restriction of $D^\ell_{m,0}$ to the 2-sphere $\Omega_2$ at $\gamma_{\v\omega}=0$ recovers (up to complex conjugation) the spherical harmonics $Y_{\ell m}$, with orthogonality relations
+\begin{align}
+  \int_{\Omega_2} \d\Omega_2\p{\v v}
+  D^{\ell}_{m,0}\p{\v v}^* D^{\ell'}_{m',0}\p{\v v}
+  = \f{4\pi}{2\ell+1} \delta_{\ell\ell'} \delta_{mm'},
+  \label{eq:ortho_sphere}
+\end{align}
+where integration is performed over all $\v v\in\Omega_2\simeq[0,2\pi]\times[0,\pi]$ with integration measure $\d\Omega_2\p{\v v} \equiv \d\alpha_{\v v} \d\beta_{\v v} \sin\beta_{\v v}$.
+The same orthogonality relations in Eqs.~\eqref{eq:ortho_full} and \eqref{eq:ortho_sphere} are inherited by the rotation matrix elements $\D_{mn}^\ell$ through the fact $\R\p{\v\omega}$ is the adjoint action of $R\p{\v\omega}$ (see Appendix \ref{sec:ortho}), and the functions $\set{\D_{m,0}^\ell}$ restricted to $\Omega_2$ are similarly another orthogonal basis for the space $L^2\p{\Omega_2}$ of square-integrable functions on the sphere.
 
@@
 We are thus nominally restricted to measuring projectors $\Pi_{\v v\mu} \equiv \op{\mu_{\v v}}$, where $\ket{\mu_{\v v}}\equiv R\p{\v v}\ket{\mu}$ is a state with spin projection $\mu$ onto a measurement axis determined by a point $\v v$ on the 2-sphere $\Omega_2$.
-For any fixed axis $\v v$, the sets $\set{\Pi_{\v v\mu}}$ and $\set{T_{\v v\ell,0}}$ are both orthonormal bases for the space of diagonal operators in the basis $\set{\ket{\mu_{\v v}}}$.
+For any fixed axis $\v v$, the sets $\set{\Pi_{\v v\mu}}$ and $\set{T_{\v v\ell,0}}$ are both orthonormal bases for the space of operators that are diagonal in the basis $\set{\ket{\mu_{\v v}}}$.
 By a simple change of operator basis, measuring the projectors $\set{\Pi_{\v v\mu}}$ is therefore equivalent to measuring the transition operators $\set{T_{\v v\ell,0}}$, and provides data on the expectation values $\bk{T_{\v \ell,0}}_\rho\equiv\tr\p{\rho T_{\v v\ell,0}}$.
@@
 \end{align}
-where the orthogonality of the transition operators motivates decomposing $C_{\ell mk}\p{\v v}=\delta_{k\ell}C_{\ell m}\p{\v v}$, which in turn implies that
+where the orthogonality of the transition operators motivates the ansatz $C_{\ell mk}\p{\v v}=\delta_{k\ell}C_{\ell m}\p{\v v}$, which in turn implies that
 \begin{align}
@@
 for all $\ell$.
-In fact, the problem of finding suitable axes $V$ and coefficients $C_{\ell m}\p{\v v}$ to satisfy Eq.~\eqref{eq:tomo_recovery}, solving which would enable the recovery of arbitrary qudit states through Eqs.~\eqref{eq:trans_state} and \eqref{eq:state_recon}, can be mapped onto the well-studied problem of signal recovery on the sphere \cite{driscoll1994computing, mcewen2011novel, mcewen2011sampling, khalid2014optimaldimensionality, rauhut2011sparse, alem2012sparse}.
+In fact, the problem of finding suitable axes $V$ and coefficients $C_{\ell m}\p{\v v}$ to satisfy Eq.~\eqref{eq:tomo_recovery}, solving which would enable the recovery of arbitrary spin qudit states, can be mapped onto the well-studied problem of signal recovery on the sphere \cite{mcewen2011novel, mcewen2011sampling, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
 The signal recovery problem can be stated as follows: given a square-integrable function $f\in L^2\p{\Omega_2}$ on the sphere, with the harmonic expansion
@@
 \end{align}
-which is a stronger version of the condition that we found for the tomography problem in Eq.~\eqref{eq:tomo_recovery}.
-If, as with the spherical harmonics $Y_{\ell m}$, the functions $B_{\ell m}$ are indexed by an integer degree $\ell\ge0$ that is preserved under rotations of the sphere (which automatically implies integer orders $\abs{m}\le\ell$), and the function $f$ is {\it band-limited} in the sense that $f_{\ell m}=0$ for all $\ell\ge L$, then the sampling problem in Eq.~\eqref{eq:full_recovery} is provably solvable with a suitable choice of $\abs{V}=L^2$ points \cite{freeden2008spherical}.
-A solution to the classical signal recovery problem Eq.~\eqref{eq:full_recovery} with $B_{\ell m}=\D^\ell_{m,0}$ automatically solves the quantum tomography problem in Eq.~\eqref{eq:tomo_recovery}, which implies the existence of $d^2$ measurement axes that suffice to reconstruct arbitrary states of $d$-level spin qudit, which are essentially ``band-limited'' at degree $\ell<d$.
+which is a stronger version of the condition that we found for the spin qudit tomography problem in Eq.~\eqref{eq:tomo_recovery}.
+We will refer to Eqs.~\eqref{eq:full_recovery} and \eqref{eq:tomo_recovery} respectively as the {\it full recovery problem} and {\it reduced recovery problem}.
+Any solution to the full recovery problem with basis functions $B_{\ell m}=\D^\ell_{m,0}$ automatically solves the reduced recovery problem.
 
-Moreover, for any fixed degree $\ell$, the problem in Eq.~\eqref{eq:tomo_recovery} is equivalent to recovery of a function in the degree-$\ell$ subspace of $L^2\p{\Omega_2}$, which is possible with $\abs{V}=2\ell+1$ samples (equal to the dimension of the degree-$\ell$ subspace) \cite{freeden2008spherical}.
-In our case, the degree $\ell$ takes a maximal value of $\ell_{\t{max}}=d-1$, so state recovery requires at least as many measurement axes as there are transition operators with degree $\ell_{\t{max}}$, namely $2\ell_{\t{max}}+1=2d-1$.
+If
+\begin{enumerate*}
+\item the functions $B_{\ell m}$ are indexed by an integer degree $\ell\ge0$ that is preserved under rotations of the sphere, as with the spherical harmonics $Y_{\ell m}$ or $\D^\ell_{m,0}$, and
+\item the function $f$ is {\it band-limited} at degree $L$, which is to say that $f_{\ell m}=0$ for all $\ell\ge L$,
+\end{enumerate*}
+then the full recovery problem in Eq.~\eqref{eq:full_recovery} is provably solvable with a suitable choice of $\abs{V}=L^2$ points \cite{freeden2008spherical, freeden2018spherical}.
+The existence of these solutions to the full recovery problem in turn implies the existence of $d^2$ measurement axes that suffice to reconstruct arbitrary states of $d$-level spin qudit, whose possible states are essentially ``band-limited'' at degree $d$.
 
-[\red{Note: finishing comments about why one could expect more than $2d-1$ axes to be necessary, and what conditions are needed for $2d-1$ to be sufficient.}]
+Moreover, any solution to the full recovery problem with a particular set of basis functions, such the spherical harmonics $Y_{\ell m}$, can be readily translated into a solution with a different set of basis functions, such as the adjoint rotation matrix elements $\D^\ell_{m,0}$.
+This translation enables importing a host of existing signal recovery algorithms \cite{mcewen2011novel, mcewen2011sampling, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality} for the task of spin qudit tomography.
+Specifically, given a set of points $V\subset\Omega_2$ and coefficients $C_{\ell m}\p{\v v}$ that satisfy Eq.~\eqref{eq:full_recovery} with basis functions $Y_{\ell m}$, we can construct new coefficients $\tilde C_{\ell m}\p{\v v}$ that satisfy the same condition with basis functions $\D^\ell_{m,0}$ via
+\begin{align}
+  \tilde C_{\ell m} \equiv \sum_n G^\ell_{mn} C_{\ell n},
+\end{align}
+where
+\begin{align}
+  G^\ell_{mn} \equiv \f{2\ell+1}{4\pi}
+  \int_{\Omega_2} \d\Omega_2\p{\v v}
+  \D^\ell_{m,0}\p{\v v}^* Y_{\ell n}\p{\v v},
+\end{align}
+is an inner product of the functions $\D^\ell_{m,0}$ and $Y_{\ell n}$ [\red{does this need to be derived in an Appendix?}].
 
-\vspace{3cm}
+Finally, for any fixed degree $\ell$, finding solutions to the reduced recovery problem in Eq.~\eqref{eq:tomo_recovery} is equivalent to recovery of a function in the degree-$\ell$ subspace of $L^2\p{\Omega_2}$, which is possible with $\abs{V}=2\ell+1$ samples (equal to the dimension of the degree-$\ell$ subspace) \cite{freeden2008spherical}.
+In our case, the degree $\ell$ takes a maximal value of $\ell_{\t{max}}\equiv d-1$, so state recovery requires at least as many measurement axes as there are transition operators with degree $\ell_{\t{max}}$, namely $2\ell_{\t{max}}+1=2d-1$.
+Since
+\begin{enumerate*}
+\item there are only $2\ell+1\le2d-1$ transition operators of degree $\ell\le\ell_{\t{max}}$, and
+\item measuring spin projections along any axis simultaneously provides measurement data for transition operators of all degrees $\ell\le\ell_{\t{max}}$,
+\end{enumerate*}
+one might expect that $2d-1$ suitably chosen measurement axes would always suffice to perform full tomography of a $d$-level spin qudit.
+Unfortunately, we have no general proof of this conjecture, which may require a deeper analysis of fixed-degree signal recovery on the sphere.
+Nonetheless, in the following section we show by example that $2d-1$ measurement axes suffice for all $d\le20$.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Tomography protocol and numerical experiments}
+\section{Tomographic protocol and numerical experiments}
+
+[\red{Mention sparse signal recovery protocols \cite{rauhut2011sparse, alem2012sparse}.}]
+
+[\red{Think about maximizing the determinant of the coefficient matrix $C_{\ell m}\p{\v v}$?}]
 
 [\red{Note: the text below is copy/pasted from a different set of notes.  I still need to adapt this text for this paper.}]
-
-[\red{Think about maximizing the determinant of the coefficient matrix $C_{m\v v}^\ell$.}]
 
@@
 \begin{align}
-  T_{\ell m} = \sqrt{\f{2\ell+1}{2s+1}} \sum_{\mu,\nu=-s}^s
+  T_{\ell m} \equiv \sqrt{\f{2\ell+1}{2s+1}} \sum_{\mu,\nu=-s}^s
   \bk{s\mu;\ell m|s\nu} \op{\nu}{\mu},
@@
 \end{align}
-Replacing the Clebsch-Gordan coefficients by the Wigner 3-$j$ symbols with the identity
+Replacing Clebsch-Gordan coefficients by Wigner 3-$j$ symbols with the identity
 \begin{align}
@@
 %%% End:
+
+signal recovery literature:
+-- a $\sim2d^2$ signaling theorem \cite{mcewen2011novel}
+-- a nice review of these methods \cite{mcewen2011sampling}
+-- an empirically accurate (albeit without rigorous proofs/guarantees) $d^2$ sampling method \cite{khalid2014optimaldimensionality}
+-- current status of the theory behind sparse signal recovery with random sampling \cite{rauhut2011sparse} (note: existing bounds are not believed to be tight)
+-- nice numerical method for sparse signal recovery \cite{alem2012sparse}
+-- useful textbook references \cite{freeden2008spherical, freeden2018spherical}

### 2020-10-20 sun_phases/spin_qudit_tomo.tex
added statement of 2d-1 conjecture in the language of classical signal recovery

@@
-\documentclass[nofootinbib,notitlepage,twocolumn]{revtex4-2}
+\documentclass[notitlepage,twocolumn]{revtex4-2}
 
@@
 one might expect that $2d-1$ suitably chosen measurement axes would always suffice to perform full tomography of a $d$-level spin qudit.
+In the language of classical signal recovery, we would expect there to exist a set of $2\ell_{\t{max}}+1$ points on the sphere that enable the recovery of any degree-$\ell$ function with $\ell\le\ell_{\t{max}}$ (given prior knowledge of $\ell$) \footnote{We could further conjecture that if {\it any} set of points $V\subset\Omega_2$ enables the recovery of degree-$\ell_{\t{max}}$ functions on the sphere, then $V$ would allow recovery degree-$\ell$ functions with $\ell\le\ell_{\t{max}}$.
+However, this condition is stronger than that necessary to guarantee full recovery of a $d$-level spin qudit state with $2d-1$ measurement axes.}.
 Unfortunately, we have no general proof of this conjecture, which may require a deeper analysis of fixed-degree signal recovery on the sphere.

### 2020-10-20 sun_phases/spin_qudit_tomo.tex
added proof of orthogonality relations on the sphere

@@
 \end{align}
-is an inner product of the functions $\D^\ell_{m,0}$ and $Y_{\ell n}$ [\red{does this need to be derived in an Appendix?}].
+is an inner product of the functions $\D^\ell_{m,0}$ and $Y_{\ell n}$ [\red{does this need to be derived in its own Appendix?}].
 
@@
   \delta_{\ell\ell'} \delta_{mm'} \delta_{nn'},
+  \label{eq:ortho_full_apndx}
 \end{align}
@@
 This property of $\D^\ell$ follows from the fact that it is an adjoint representation of SO(3), which is isomorphic to the standard representation $D^\ell$ \cite{hall2015lie}.
+
+The decomposition $D^\ell_{mn}\p{\v\omega} = e^{-\i\alpha_{\v\omega}m} d^\ell_{mn}\p{\beta_{\v\omega}} e^{-\i\gamma_{\v\omega}n}$, where $d^\ell_{mn}\p{\beta}\equiv D^\ell_{mn}\p{0,\beta,0}$ is
+a matrix element of a reduced fundamental (Wigner) rotation matrix, implies that setting $n'=n$ in in Eq.~\eqref{eq:ortho_full_apndx} makes the integrand independent of $\gamma_{\v\omega}$.
+Evaluating the integral over $\gamma_{\v\omega}$ and dividing through by $2\pi$ then recovers the orthogonality relations
+\begin{align}
+  \int_{\Omega_2} \d\Omega_2\p{\v v}
+  D^{\ell}_{mn}\p{\v v}^* D^{\ell'}_{m'n}\p{\v v}
+  = \f{4\pi}{2\ell+1} \delta_{\ell\ell'} \delta_{mm'},
+\end{align}
+where integration is performed over all $\v v\in\Omega_2\simeq[0,2\pi]\times[0,\pi]$ with integration measure $\d\Omega_2\p{\v v} \equiv \d\alpha_{\v v} \d\beta_{\v v} \sin\beta_{\v v}$.
+A similar decomposition of the adjoint rotation matrix elements $\D^\ell_{mn}\p{\v\omega}$ (see Appendix \ref{sec:rotations}) implies the identical orthogonality relations.
 

### 2020-10-20 sun_phases/spin_qudit_tomo.tex
added a tiny bit to the intro added an appendix deriving the translation between solutions to the signal recovery problem in different bases

@@
   We then provide an explicit optimized tomographic protocol, as well as numerical experiments to show that $r_d=2d-1$ for $d\le20$.
-  [\red{Note: there are also a few other things that I want to look into, such as (1) adaptive tomography protocols, (2) proofs about something like the number of measurements necessary to reconstruct an arbitrary state to a given precision, (3) a simplified tomography procedure for measuring pure states, (4) the trade-off between measurement precision and the \# of measurement axes (for a fixed total number of measurements).]}
+  [\red{Note: there are also a few other things that I want to look into, including (1) a simplified tomography procedure for measuring pure states, (2) trade-offs between measurement precision and the \# of measurement axes (for a fixed total number of measurements; these will likely be numerical experiments), (3) proofs about something like the number of measurements necessary to reconstruct an arbitrary state to a given precision, and (4) adaptive tomography protocols.]}
 \end{abstract}
@@
 
-In this work, we provide a new and simple framework for performing spin qudit tomography.
-We begin by mapping this quantum tomography problem onto the well-studied classical problem of signal recovery on the sphere, thereby importing a host of existing literature and machinery that has been developed for this task \cite{mcewen2011novel, mcewen2011sampling, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
+In this work, we provide an efficient framework for performing spin qudit tomography using only measurements of spin projection onto different spatial axes.
+We first introduce, in Section \ref{sec:transition_ops}, a basis of operators that have a simple physical interpretation, and transform nicely under rotations in 3D space, thereby respecting the symmetries of the tomography task at hand.
+We then map this quantum tomography problem onto the well-studied classical problem of signal recovery on the sphere in Section \ref{sec:signal_recovery}, thereby importing a host of existing literature and machinery that has been developed for this task \cite{mcewen2011novel, mcewen2011sampling, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
+In particular, this mapping allows us to bound the minimum number of measurement axes necessary to perform full tomography on a $d$-level spin qudit, $r_d$, by $2d-1\le r_d\le d^2$.
+We provide an example of an optimized tomography protocol in Section \ref{sec:protocol}, as well as numerical experiments to show that $r_d=2d-1$ for $d\le20$.
 
-[\red{finish intro, provide outline of the paper}]
+[\red{finish intro}]
 
@@
 \section{Spin transition operators and rotation matrices}
+\label{sec:transition_ops}
 
@@
 \section{Spin tomography as signal recovery on the sphere}
+\label{sec:signal_recovery}
 
@@
 For any fixed axis $\v v$, the sets $\set{\Pi_{\v v\mu}}$ and $\set{T_{\v v\ell,0}}$ are both orthonormal bases for the space of operators that are diagonal in the basis $\set{\ket{\mu_{\v v}}}$.
-By a simple change of operator basis, measuring the projectors $\set{\Pi_{\v v\mu}}$ is therefore equivalent to measuring the transition operators $\set{T_{\v v\ell,0}}$, and provides data on the expectation values $\bk{T_{\v \ell,0}}_\rho\equiv\tr\p{\rho T_{\v v\ell,0}}$.
+By a simple change of operator basis, measuring the projectors $\set{\Pi_{\v v\mu}}$ is therefore equivalent to measuring the transition operators $\set{T_{\v v\ell,0}}$, and provides data on the expectation values $\bk{T_{\v \ell,0}}_\rho$, where $\bk{X}_\rho\equiv\tr\p{\rho X}$.
 
@@
 \end{align}
-is an inner product of the functions $\D^\ell_{m,0}$ and $Y_{\ell n}$ [\red{does this need to be derived in its own Appendix?}].
+is an inner product of the functions $\D^\ell_{m,0}$ and $Y_{\ell n}$ (see Appendix \ref{sec:basis_change}).
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Tomographic protocol and numerical experiments}
+\section{Tomography protocol and numerical experiments}
+\label{sec:protocol}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Changing bases for the signal recovery problem}
+\label{sec:basis_change}
+
+Here we show how one solution to the problem of signal recovery on the sphere with a particular basis of functions (such as the spherical harmonics $Y_{\ell m}$) can be translated into a solution with a different basis (such as the adjoint rotation matrix elements $\D^\ell_{m,0}$).
+Consider a square-integrable function on the sphere, $f\in L^2\p{\Omega_2}$, with the harmonic expansion
+\begin{align}
+  f\p{\v v} = \sum_{\ell,m} f_{\ell m} B_{\ell m},
+\end{align}
+where $f_{\ell m}\in\CC$ is a coefficient and $\set{B_{\ell m}}$ is a complete basis for $L^2\p{\Omega_2}$.
+Without loss of generality, we will assume that for a fixed degree $\ell$ the functions $\set{B_{\ell m}}$ span the degree-$\ell$ subspace of $L^2\p{\Omega_2}$, and that these functions satisfy orthogonality relations analogous to that in Eq.~\eqref{eq:ortho_sphere}.
+A solution to the signal recovery problem amounts to a set of points $V=\set{\v v}\subset\Omega_2$ and coefficients $C_{\ell m}\p{\v v}$ for which
+\begin{align}
+  f_{\ell m} = \sum_{\v v} C_{\ell m}\p{\v v} f\p{\v v}.
+  \label{eq:recovery_solution}
+\end{align}
+Consider now a different basis $\set{\tilde B_{\ell m}}$ for $L^2\p{\Omega_2}$, with which
+\begin{align}
+  f\p{\v v} = \sum_{\ell, m} \tilde f_{\ell m} \tilde B_{\ell m},
+\end{align}
+where we make the same general assumptions about $\set{\tilde B_{\ell m}}$ that we made for $\set{B_{\ell m}}$.
+Substituting the harmonic expansion of $f\p{\v v}$ in the original basis $\set{B_{\ell m}}$, we find that
+\begin{align}
+  \tilde f_{\ell m} = \int_{\Omega_2} \d\Omega_2\p{\v v}
+  \tilde B_{\ell m}\p{\v v}^* f\p{\v v}
+  = \sum_n G^\ell_{mn} f_{\ell n},
+  &&
+  G^\ell_{mn} \equiv \int_{\Omega_2} \d\Omega_2\p{\v v}
+  \tilde B_{\ell m}\p{\v v}^* B_{\ell n}\p{\v v},
+  \label{eq:basis_change}
+\end{align}
+where integration is performed over all $\v v\in\Omega_2\simeq[0,2\pi]\times[0,\pi]$ with integration measure $\d\Omega_2\p{\v v} \equiv \d\alpha_{\v v} \d\beta_{\v v} \sin\beta_{\v v}$.
+Note that the inner product between functions $\tilde B_{\ell m}\p{\v v}$ and $B_{kn}\p{\v v}$ for $k\ne\ell$ vanishes because these functions live in orthogonal subspaces of $L^2\p{\Omega_2}$ \cite{freeden2008spherical}.
+Substituting Eq.~\eqref{eq:recovery_solution} into Eq.~\eqref{eq:basis_change}, we find that
+\begin{align}
+  \tilde f_{\ell m}
+  = \sum_n G^\ell_{mn} \sum_{\v v} C_{\ell n}\p{\v v} f\p{\v v}
+  = \sum_{\v v} \tilde C_{\ell m}\p{\v v} f\p{\v v},
+  &&
+  \tilde C_{\ell m}\p{\v v} \equiv \sum_n G^\ell_{mn} C_{\ell n}\p{\v v}.
+\end{align}
+The coefficients $\tilde C_{\ell m}\p{\v v}$ thus solve the signal recovery problem with the new basis $\set{\tilde B_{\ell m}}$ for $L^2\p{\Omega_2}$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Transition operator product expansion}

### 2020-10-21 sun_phases/spin_qudit_tomo.tex
aesthetic changes to text and notation

@@
 }
-
-%%% header / footer
-\usepackage{fancyhdr} % easier header and footer management
-\pagestyle{fancy} % page formatting style
-\fancyhf{} % clear all header and footer text
-\renewcommand{\headrulewidth}{0pt} % remove horizontal line in header
-\usepackage{lastpage} % for referencing last page
-\cfoot{\thepage~of \pageref{LastPage}} % "x of y" page labeling
 
@@
 \end{align}
-where $f_{\ell m}$ are expansion coefficients and $\set{B_{\ell m}}$ is an orthogonal basis for $L^2\p{\Omega_2}$, find a set of points $V=\set{\v v}$ and associated coefficients $C_{\ell m}\p{\v v}$ with which we can reconstruct $f$, or equivalently its coefficients $f_{\ell m}$, from knowledge of the function's value $f\p{\v v}$ at all points $\v v\in V$; that is
+where $f_{\ell m}$ are expansion coefficients and $\set{B_{\ell m}}$ is an orthogonal basis for $L^2\p{\Omega_2}$, find a set of points $V=\set{\v v}\subset\Omega_2$ and associated coefficients $C_{\ell m}\p{\v v}$ with which we can reconstruct $f$, or equivalently its coefficients $f_{\ell m}$, from knowledge of the function's value $f\p{\v v}$ at all points $\v v\in V$; that is
 \begin{align}
@@
 \begin{enumerate*}
-\item the functions $B_{\ell m}$ are indexed by an integer degree $\ell\ge0$ that is preserved under rotations of the sphere, as with the spherical harmonics $Y_{\ell m}$ or $\D^\ell_{m,0}$, and
+\item the functions $B_{\ell m}$ are indexed by an integer degree $\ell\ge0$ that is preserved under rotations of the sphere and an integer order $m\in\set{-\ell,-\ell-1,\cdots,\ell}$, as with the spherical harmonics $Y_{\ell m}$ and $\D^\ell_{m,0}$, and
 \item the function $f$ is {\it band-limited} at degree $L$, which is to say that $f_{\ell m}=0$ for all $\ell\ge L$,
@@
 \end{align}
-is an inner product of the functions $\D^\ell_{m,0}$ and $Y_{\ell n}$ (see Appendix \ref{sec:basis_change}).
+is an inner product of $\D^\ell_{m,0}$ and $Y_{\ell n}$ (see Appendix \ref{sec:basis_change}).
 
 Finally, for any fixed degree $\ell$, finding solutions to the reduced recovery problem in Eq.~\eqref{eq:tomo_recovery} is equivalent to recovery of a function in the degree-$\ell$ subspace of $L^2\p{\Omega_2}$, which is possible with $\abs{V}=2\ell+1$ samples (equal to the dimension of the degree-$\ell$ subspace) \cite{freeden2008spherical}.
-In our case, the degree $\ell$ takes a maximal value of $\ell_{\t{max}}\equiv d-1$, so state recovery requires at least as many measurement axes as there are transition operators with degree $\ell_{\t{max}}$, namely $2\ell_{\t{max}}+1=2d-1$.
+In the case of spin qudit tomography, the degree $\ell$ takes a maximal value of $\ell_{\t{max}}\equiv d-1$, so state recovery requires at least as many measurement axes as there are transition operators with degree $\ell_{\t{max}}$, namely $2\ell_{\t{max}}+1=2d-1$.
 Since
@@
 
-As functions of $\v\omega$, the fundamental rotation matrix elements $D_{mn}^\ell$ and satisfy the orthogonality relations \cite{brown2003rotational}
+As functions of Euler angle triplets $\v\omega=\p{\alpha_{\v\omega},\beta_{\v\omega},\gamma_{\v\omega}}$, the fundamental (Wigner) rotation matrix elements $D_{mn}^\ell$ and satisfy the orthogonality relations \cite{brown2003rotational}
 \begin{align}
@@
 The decomposition $D^\ell_{mn}\p{\v\omega} = e^{-\i\alpha_{\v\omega}m} d^\ell_{mn}\p{\beta_{\v\omega}} e^{-\i\gamma_{\v\omega}n}$, where $d^\ell_{mn}\p{\beta}\equiv D^\ell_{mn}\p{0,\beta,0}$ is
-a matrix element of a reduced fundamental (Wigner) rotation matrix, implies that setting $n'=n$ in in Eq.~\eqref{eq:ortho_full_apndx} makes the integrand independent of $\gamma_{\v\omega}$.
+a matrix element of a reduced fundamental (Wigner) rotation matrix \cite{rose1957elementary}, implies that setting $n'=n$ in in Eq.~\eqref{eq:ortho_full_apndx} makes the integrand independent of $\gamma_{\v\omega}$.
 Evaluating the integral over $\gamma_{\v\omega}$ and dividing through by $2\pi$ then recovers the orthogonality relations
@@
 \begin{align}
-  \bk{\ell_1 m_1; \ell_2 m_2| \ell_3 m_3}
-  &= \p{-1}^{\ell_2+m_2} \sqrt{\f{2\ell_3+1}{2\ell_1+1}}
-  \bk{\ell_3,-m_3; \ell_2 m_2| \ell_1,-m_1} \\
-  \bk{\ell_1 m_1; \ell_2 m_2| \ell_3 m_3}
-  &= \p{-1}^{\ell_1+\ell_2-\ell_3}
-  \bk{\ell_1,-m_1; \ell_2,-m_2| \ell_3,-m_3},
+  \bk{\ell_1 m_1; \ell_2 m_2| L M}
+  &= \p{-1}^{\ell_2+m_2} \sqrt{\f{2L+1}{2\ell_1+1}}
+  \bk{L,-M; \ell_2 m_2| \ell_1,-m_1} \\
+  \bk{\ell_1 m_1; \ell_2 m_2| L M}
+  &= \p{-1}^{\ell_1+\ell_2-L}
+  \bk{\ell_1,-m_1; \ell_2,-m_2| L,-M},
 \end{align}
@@
 \begin{align}
-  \bk{\ell_1 m_1; \ell_2 m_2| \ell_3 m_3}
-  = \p{-1}^{2\ell_2} \p{-1}^{\ell_3-m_3} \sqrt{\p{2\ell_3+1}}
+  \bk{\ell_1 m_1; \ell_2 m_2| L M}
+  = \p{-1}^{2\ell_2} \p{-1}^{L-M} \sqrt{2L+1}
   \begin{pmatrix}
-    \ell_3 & \ell_2 & \ell_1 \\
-    -m_3 & m_2 & m_1
+    L & \ell_2 & \ell_1 \\
+    -M & m_2 & m_1
   \end{pmatrix},

### 2020-10-21 sun_phases/spin_qudit_tomo.tex
rearrange some existing text

@@
 
-\keywords{spin qudits; quantum state tomography}
+\keywords{qudits; spin qudits; quantum state tomography}
 
@@
   We consider the task of performing quantum state tomography on a $d$-state spin qudit, using only measurements of spin projection onto different quantization axes.
-  By mapping this quantum tomography problem onto the classical problem of signal recovery on the sphere, we prove that full reconstruction of arbitrary mixed states is possible, and requires a minimum number of measurement axes, $r_d$, that is bounded by $2d-1\le r_d\le d^2$.
-  We then provide an explicit optimized tomographic protocol, as well as numerical experiments to show that $r_d=2d-1$ for $d\le20$.
-  [\red{Note: there are also a few other things that I want to look into, including (1) a simplified tomography procedure for measuring pure states, (2) trade-offs between measurement precision and the \# of measurement axes (for a fixed total number of measurements; these will likely be numerical experiments), (3) proofs about something like the number of measurements necessary to reconstruct an arbitrary state to a given precision, and (4) adaptive tomography protocols.]}
+  We map this quantum tomography problem onto the classical problem of signal recovery on the sphere, thereby proving that full reconstruction of arbitrary mixed states is possible and requires a minimum number of measurement axes, $r_d$, that is bounded by $2d-1\le r_d\le d^2$.
+  Our mapping comes with a straightforward prescription for converting existing signal recovery algorithms into spin qudit tomography protocols, and numerical experiments show that $r_d=2d-1$ for $d\le20$.
+  [\red{Note: there are also a few other things that I want to look into, including (1) a simplified tomography procedure for pure states, (2) trade-offs between measurement precision and the \# of measurement axes (for a fixed total number of measurements; these will likely be numerical experiments), (3) proofs about the something like the number of measurements necessary to reconstruct an arbitrary state to a given precision, and (4) adaptive tomography protocols.]}
 \end{abstract}
@@
 
-Due in part to advancements in experimental capabilities to address nuclear spins in ultracold atomic systems [\red{CITE}], as well as developments in molecular cooling techniques [\red{CITE}], a particular setting of growing interest is the spin qudit, which describes a quantum angular momentum degree of freedom.
-Multilevel spin qudits can provide advantages for quantum sensing \cite{evrard2019enhanced}, enable quantum simulations of SU($N$) magnetism \cite{banerjee2013atomic, cazalilla2014ultracold}, and offer unique capabilities for quantum error correction \cite{albert2020robust, gross2020encoding}.
+Due in part to advancements in experimental capabilities to address nuclear spins in ultracold atomic systems [\red{CITE}], as well as developments in molecular cooling techniques [\red{CITE}], a particular setting of growing interest is the spin qudit, or a multilevel quantum angular momentum degree of freedom.
+Spin qudits can provide advantages over their qubit counterparts for quantum sensing \cite{evrard2019enhanced}, enable quantum simulations of SU($N$) magnetism \cite{banerjee2013atomic, cazalilla2014ultracold}, and offer unique capabilities for quantum error correction \cite{albert2020robust, gross2020encoding}.
 In all cases, quantum state tomography is necessary to take full advantage of a spin qudit.
@@
 In particular, this mapping allows us to bound the minimum number of measurement axes necessary to perform full tomography on a $d$-level spin qudit, $r_d$, by $2d-1\le r_d\le d^2$.
-We provide an example of an optimized tomography protocol in Section \ref{sec:protocol}, as well as numerical experiments to show that $r_d=2d-1$ for $d\le20$.
+We provide an example of an optimized tomography protocol in Section \ref{sec:protocol} [\red{I'll probably use modified versions of existing signal recovery algorithms rather than a new tomography protocol}], as well as numerical experiments to show that $r_d=2d-1$ for $d\le20$.
 
@@
 
-Moreover, any solution to the full recovery problem with a particular set of basis functions, such the spherical harmonics $Y_{\ell m}$, can be readily translated into a solution with a different set of basis functions, such as the adjoint rotation matrix elements $\D^\ell_{m,0}$.
+Moreover, any solution to the full recovery problem with a particular set of basis functions, such the spherical harmonics $Y_{\ell m}$, can be readily translated into a solution with a different set of basis functions, such as the adjoint rotation matrix elements $\D^\ell_{m,0}$ (see Appendix \ref{sec:basis_change}).
 This translation enables importing a host of existing signal recovery algorithms \cite{mcewen2011novel, mcewen2011sampling, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality} for the task of spin qudit tomography.
@@
 \end{align}
-is an inner product of $\D^\ell_{m,0}$ and $Y_{\ell n}$ (see Appendix \ref{sec:basis_change}).
+is an inner product of $\D^\ell_{m,0}$ and $Y_{\ell n}$.
 
@@
 \item there are only $2\ell+1\le2d-1$ transition operators of degree $\ell\le\ell_{\t{max}}$, and
-\item measuring spin projections along any axis simultaneously provides measurement data for transition operators of all degrees $\ell\le\ell_{\t{max}}$,
+\item measuring spin projections along any axis simultaneously provides measurement data for transition operators of {\it all} degrees $\ell\le\ell_{\t{max}}$,
 \end{enumerate*}

### 2020-10-22 sun_phases/spin_qudit_tomo.tex
minor change to appendix

@@
   We consider the task of performing quantum state tomography on a $d$-state spin qudit, using only measurements of spin projection onto different quantization axes.
-  We map this quantum tomography problem onto the classical problem of signal recovery on the sphere, thereby proving that full reconstruction of arbitrary mixed states is possible and requires a minimum number of measurement axes, $r_d$, that is bounded by $2d-1\le r_d\le d^2$.
-  Our mapping comes with a straightforward prescription for converting existing signal recovery algorithms into spin qudit tomography protocols, and numerical experiments show that $r_d=2d-1$ for $d\le20$.
+  By mapping this quantum tomography problem onto the classical problem of signal recovery on the sphere, we prove that full reconstruction of arbitrary mixed states is possible, and requires a minimum number of measurement axes, $r_d$, that is bounded by $2d-1\le r_d\le d^2$.
+  We provide explicit optimized tomography protocols, and show that $r_d=2d-1$ for $d\le20$ with numerical experiments.
   [\red{Note: there are also a few other things that I want to look into, including (1) a simplified tomography procedure for pure states, (2) trade-offs between measurement precision and the \# of measurement axes (for a fixed total number of measurements; these will likely be numerical experiments), (3) proofs about the something like the number of measurements necessary to reconstruct an arbitrary state to a given precision, and (4) adaptive tomography protocols.]}

### 2020-10-22 sun_phases/spin_qudit_tomo.tex
discuss different random sampling schemes

@@
 Due in part to advancements in experimental capabilities to address nuclear spins in ultracold atomic systems [\red{CITE}], as well as developments in molecular cooling techniques [\red{CITE}], a particular setting of growing interest is the spin qudit, or a multilevel quantum angular momentum degree of freedom.
-Spin qudits can provide advantages over their qubit counterparts for quantum sensing \cite{evrard2019enhanced}, enable quantum simulations of SU($N$) magnetism \cite{banerjee2013atomic, cazalilla2014ultracold}, and offer unique capabilities for quantum error correction \cite{albert2020robust, gross2020encoding}.
+Spin qudits can provide advantages over their qubit counterparts for quantum sensing \cite{evrard2019enhanced}, enable quantum simulations of SU($d$) magnetism \cite{banerjee2013atomic, cazalilla2014ultracold}, and offer unique capabilities for quantum error correction \cite{albert2020robust, gross2020encoding}.
 In all cases, quantum state tomography is necessary to take full advantage of a spin qudit.
@@
 We first introduce, in Section \ref{sec:transition_ops}, a basis of operators that have a simple physical interpretation, and transform nicely under rotations in 3D space, thereby respecting the symmetries of the tomography task at hand.
-We then map this quantum tomography problem onto the well-studied classical problem of signal recovery on the sphere in Section \ref{sec:signal_recovery}, thereby importing a host of existing literature and machinery that has been developed for this task \cite{mcewen2011novel, mcewen2011sampling, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
+We then map this quantum tomography problem onto the well-studied classical problem of signal recovery on the sphere in Section \ref{sec:signal_recovery}, thereby importing a host of existing literature and mathematical machinery that has been developed for this task \cite{mcewen2011novel, mcewen2011sampling, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
 In particular, this mapping allows us to bound the minimum number of measurement axes necessary to perform full tomography on a $d$-level spin qudit, $r_d$, by $2d-1\le r_d\le d^2$.
 We provide an example of an optimized tomography protocol in Section \ref{sec:protocol} [\red{I'll probably use modified versions of existing signal recovery algorithms rather than a new tomography protocol}], as well as numerical experiments to show that $r_d=2d-1$ for $d\le20$.
+
+We generally defer mathematical details and derivations to the appendices, keeping the main text focused on a high-level discussion of our results.
 
@@
 The adjoint rotation matrix elements $\D_{mn}^\ell\p{\v\omega}\equiv\obk{T_{\ell m}|\R\p{\v\omega}|T_{\ell n}}$ are an analog of the fundamental (Wigner) rotation matrix elements $D_{mn}^\ell\p{\v\omega}\equiv\bk{\ell m|R\p{\v\omega}|\ell n}$ that govern rotations of spin-$\ell$ particles and spherical harmonics $Y_{\ell m}$, where $\ket{\ell m}$ denotes the state of a spin-$\ell$ qudit with spin projection $m$ onto a quantization axis.
-Throughout this work, we will primarily consider rotations determined by points on the 2-sphere $\Omega_2$ at $\gamma_{\v\omega}=0$.
-For ease of notation, we therefore implicitly associate doublets $\v v=\p{\alpha_{\v v},\beta_{\v v}}$ with the triplet $\p{\alpha_{\v v},\beta_{\v v},0}$, so that for example $R\p{\v v} = R\p{\alpha_{\v v},\beta_{\v v},0}$ and $T_{\v v\ell m} = T_{\p{\alpha_{\v v},\beta_{\v v},0},\ell m}$.
 
+It will be useful to think of rotation matrix elements $D^\ell_{mn}$ and $\D^\ell_{mn}$ as scalar functions of Euler angles $\v\omega\in\Omega_3$.
 As functions of $\v\omega$, the fundamental rotation matrix elements $D_{mn}^\ell$ satisfy the orthogonality relations \cite{brown2003rotational}
@@
 where integration is performed over all $\v\omega\in\Omega_3\simeq[0,2\pi]\times[0,\pi]\times[0,2\pi]$ with integration measure $\d\Omega_3\p{\v\omega} \equiv \d\alpha_{\v\omega} \d\beta_{\v\omega} \d\gamma_{\v\omega} \sin\beta_{\v\omega}$.
-In particular, the restriction of $D^\ell_{m,0}$ to the 2-sphere $\Omega_2$ at $\gamma_{\v\omega}=0$ recovers (up to complex conjugation) the spherical harmonics $Y_{\ell m}$, with orthogonality relations
+In particular, the restriction of $D^\ell_{m,0}$ to the sphere $\Omega_2$ at $\gamma_{\v\omega}=0$ recovers (up to complex conjugation) the spherical harmonics $Y_{\ell m}$, with orthogonality relations
 \begin{align}
@@
 \end{align}
-where integration is performed over all $\v v\in\Omega_2\simeq[0,2\pi]\times[0,\pi]$ with integration measure $\d\Omega_2\p{\v v} \equiv \d\alpha_{\v v} \d\beta_{\v v} \sin\beta_{\v v}$.
+where integration is performed over all $\v v\in\Omega_2\simeq[0,2\pi]\times[0,\pi]$ with integration measure $\d\Omega_2\p{\v v} \equiv \d\alpha_{\v v} \d\beta_{\v v} \sin\beta_{\v v}$, and for ease of notation we implicitly associate doublets $\v v=\p{\alpha_{\v v},\beta_{\v v}}$ with the triplet $\p{\alpha_{\v v},\beta_{\v v},0}$, so that for example $R\p{\v v} = R\p{\alpha_{\v v},\beta_{\v v},0}$ and $T_{\v v\ell m} = T_{\p{\alpha_{\v v},\beta_{\v v},0},\ell m}$.
 The same orthogonality relations in Eqs.~\eqref{eq:ortho_full} and \eqref{eq:ortho_sphere} are inherited by the rotation matrix elements $\D_{mn}^\ell$ through the fact $\R\p{\v\omega}$ is the adjoint action of $R\p{\v\omega}$ (see Appendix \ref{sec:ortho}), and the functions $\set{\D_{m,0}^\ell}$ restricted to $\Omega_2$ are similarly another orthogonal basis for the space $L^2\p{\Omega_2}$ of square-integrable functions on the sphere.
+[\red{todo: make figures visualizing $Y_{\ell m}$ and $\D^\ell_{m,0}$ for comparison}]
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Spin tomography as signal recovery on the sphere}
+\section{Tomography as signal recovery on the sphere}
 \label{sec:signal_recovery}
@@
 Our goal is to reconstruct an arbitrary state of a spin qudit from measurements of spin projection onto different quantization axes.
-We are thus nominally restricted to measuring projectors $\Pi_{\v v\mu} \equiv \op{\mu_{\v v}}$, where $\ket{\mu_{\v v}}\equiv R\p{\v v}\ket{\mu}$ is a state with spin projection $\mu$ onto a measurement axis determined by a point $\v v$ on the 2-sphere $\Omega_2$.
+We are thus nominally restricted to measuring projectors $\Pi_{\v v\mu} \equiv \op{\mu_{\v v}}$, where $\ket{\mu_{\v v}}\equiv R\p{\v v}\ket{\mu}$ is a state with spin projection $\mu$ onto a measurement axis determined by a point $\v v$ on the sphere $\Omega_2$.
 For any fixed axis $\v v$, the sets $\set{\Pi_{\v v\mu}}$ and $\set{T_{\v v\ell,0}}$ are both orthonormal bases for the space of operators that are diagonal in the basis $\set{\ket{\mu_{\v v}}}$.
@@
 \end{align}
-Expanding the rotated transition operators $T_{\v vk,0}$ according to Eq.~\eqref{eq:trans_rot}, we find that the recovery condition in Eq.~\eqref{eq:state_recon} is satisfied when
+Expanding the rotated transition operators $T_{\v vk,0}$ into a sum of un-rotated transition operators, according to Eq.~\eqref{eq:trans_rot}, we find that the recovery condition in Eq.~\eqref{eq:state_recon} is satisfied when
 \begin{align}
   T_{\ell m}
-  = \sum_{\v v,k,n} C_{\ell mk}\p{\v v} \D^k_{n,0}\p{\v v} T_{kn},
+  = \sum_{\v v,k,n} C_{\ell mk}\p{\v v} \D^k_{n,0}\p{\v v} T_{kn}.
 \end{align}
-where the orthogonality of the transition operators motivates the ansatz $C_{\ell mk}\p{\v v}=\delta_{k\ell}C_{\ell m}\p{\v v}$, which in turn implies that
+Orthogonality of the transition operators motivates the ansatz $C_{\ell mk}\p{\v v}=\delta_{k\ell}C_{\ell m}\p{\v v}$, which in turn implies that
 \begin{align}
@@
 for all $\ell$.
-In fact, the problem of finding suitable axes $V$ and coefficients $C_{\ell m}\p{\v v}$ to satisfy Eq.~\eqref{eq:tomo_recovery}, solving which would enable the recovery of arbitrary spin qudit states, can be mapped onto the well-studied problem of signal recovery on the sphere \cite{mcewen2011novel, mcewen2011sampling, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
+
+In fact, the problem of finding suitable axes $V$ and coefficients $C_{\ell m}\p{\v v}$ to satisfy Eq.~\eqref{eq:tomo_recovery}, solving which would enable the recovery of arbitrary spin qudit states, can be mapped onto the well-studied problem of signal recovery on the sphere \cite{mcewen2011novel, mcewen2011sampling, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality} [\red{todo: make figure to visualize the signal recovery problem}].
 The signal recovery problem can be stated as follows: given a square-integrable function $f\in L^2\p{\Omega_2}$ on the sphere, with the harmonic expansion
@@
 which is a stronger version of the condition that we found for the spin qudit tomography problem in Eq.~\eqref{eq:tomo_recovery}.
-We will refer to Eqs.~\eqref{eq:full_recovery} and \eqref{eq:tomo_recovery} respectively as the {\it full recovery problem} and {\it reduced recovery problem}.
+We will refer to Eq.~\eqref{eq:full_recovery} as the {\it full recovery problem}, and Eq.~\eqref{eq:tomo_recovery} as the the {\it reduced recovery problem}.
 Any solution to the full recovery problem with basis functions $B_{\ell m}=\D^\ell_{m,0}$ automatically solves the reduced recovery problem.
@@
 then the full recovery problem in Eq.~\eqref{eq:full_recovery} is provably solvable with a suitable choice of $\abs{V}=L^2$ points \cite{freeden2008spherical, freeden2018spherical}.
-The existence of these solutions to the full recovery problem in turn implies the existence of $d^2$ measurement axes that suffice to reconstruct arbitrary states of $d$-level spin qudit, whose possible states are essentially ``band-limited'' at degree $d$.
+The existence of these solutions to the full recovery problem in turn implies the existence of $d^2$ measurement axes that suffice to reconstruct arbitrary states of $d$-level spin qudit, whose possible states are essentially ``band-limited'' at degree $d$ \footnote{Spin qudit states can be faithfully represented by quasi-probability distributions on a sphere.
+For a $d$-level spin, these quasi-probability distributions are band-limited at degree $d$.}.
 
@@
 
-Finally, for any fixed degree $\ell$, finding solutions to the reduced recovery problem in Eq.~\eqref{eq:tomo_recovery} is equivalent to recovery of a function in the degree-$\ell$ subspace of $L^2\p{\Omega_2}$, which is possible with $\abs{V}=2\ell+1$ samples (equal to the dimension of the degree-$\ell$ subspace) \cite{freeden2008spherical}.
+Finally, for any fixed degree $\ell$, finding solutions to the reduced recovery problem in Eq.~\eqref{eq:tomo_recovery} is equivalent to the recovery of a function in the degree-$\ell$ subspace of $L^2\p{\Omega_2}$, which is possible with $\abs{V}=2\ell+1$ samples (equal to the dimension of the degree-$\ell$ subspace) \cite{freeden2008spherical}.
 In the case of spin qudit tomography, the degree $\ell$ takes a maximal value of $\ell_{\t{max}}\equiv d-1$, so state recovery requires at least as many measurement axes as there are transition operators with degree $\ell_{\t{max}}$, namely $2\ell_{\t{max}}+1=2d-1$.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Tomography protocol and numerical experiments}
+\section{Optimized tomography protocols}
 \label{sec:protocol}
 
-[\red{Mention sparse signal recovery protocols \cite{rauhut2011sparse, alem2012sparse}.}]
+For the practically minded, the mere existence of solutions to a problem is less interesting than the exposition of a particular solution.
+On a high level, a spin qudit tomography protocol needs to
+\begin{enumerate*}
+\item select a set of measurement axes,
+  \label{tomo:select}
+\item collect measurement data on spin projection onto these axes, and then
+  \label{tomo:measure}
+\item process the collected data to reconstruct the state of the spin qudit.
+  \label{tomo:process}
+\end{enumerate*}
+Whereas step \ref{tomo:measure} can involve a host of platform-dependent technical challenges, here we discuss the steps before and after measurement data collection.
+
+We begin by discussing the selection of measurement axes $V=\set{\v v}\subset\Omega_2$.
+In the absence of analytical tools to find ``optimal'' deterministic choice of axes, we compare three randomized selection methods that are motivated by past works on sparse signal recovery \cite{rauhut2011sparse, alem2012sparse}.
+Conceptually, the simplest method of randomly choosing axes would be to choose $\abs{V}$ points from the uniform distribution on the sphere.
+To select a point $\v v=\p{\alpha_{\v v},\beta_{\v v}}$ from the uniform distribution on the sphere, we first select a point $\p{a_{\v v},z_{\v v}}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and set $\alpha_{\v v}=2\pi a_{\v v}$ and $\beta_{\v v}=\arccos\p{1-2z_{\v v}}$.
+We will call this procedure the {\it uniform} sampling method.
+
+The work in Ref.~\cite{rauhut2011sparse} proposes an alternate sampling method for sparse signal recovery on the sphere, which consists of uniformly sampling angles $\p{\alpha_{\v v},\beta_{\v v}}\in[0,2\pi]\times[0,\pi]$.
+We will refer to this procedure as the {\it biased} sampling scheme, as it places a higher weight on points near the poles of the sphere.
+For general-purpose recovery of band-limited functions $f=\sum_{\ell,m} f_{\ell m} Y_{\ell m}$ with only a small number of non-zero spherical harmonic coefficients $f_{\ell m}$, Ref.~\cite{rauhut2011sparse} found that the biased sampling scheme was more likely to result in ``good'' points that allow for accurate signal recovery.
+
+Finally, we consider the {\it spiral} sampling scheme laid out in Ref.~\cite{alem2012sparse}, which first generates a spiral of points that go from the north pole to the south, using this spiral to define a grid of points on the entire sphere, and then randomly choosing $\abs{V}$ points on this grid.
+
+[\red{todo: make figure visualizing these sampling schemes}]
+
+\vspace{3cm}
 

### 2020-10-22 sun_phases/spin_qudit_tomo.tex
add proof of error bound in an estimated state

@@
 \newcommand{\B}{\mathcal{B}}
-\newcommand{\G}{\mathcal{G}}
+\newcommand{\N}{\mathcal{N}}
 \renewcommand{\O}{\mathcal{O}}
@@
 
-\newtheorem{theorem}{Theorem}[section]
+\let\var\relax
+\DeclareMathOperator{\var}{var}
 
@@
   By mapping this quantum tomography problem onto the classical problem of signal recovery on the sphere, we prove that full reconstruction of arbitrary mixed states is possible, and requires a minimum number of measurement axes, $r_d$, that is bounded by $2d-1\le r_d\le d^2$.
-  We provide explicit optimized tomography protocols, and show that $r_d=2d-1$ for $d\le20$ with numerical experiments.
-  [\red{Note: there are also a few other things that I want to look into, including (1) a simplified tomography procedure for pure states, (2) trade-offs between measurement precision and the \# of measurement axes (for a fixed total number of measurements; these will likely be numerical experiments), (3) proofs about the something like the number of measurements necessary to reconstruct an arbitrary state to a given precision, and (4) adaptive tomography protocols.]}
+  We conjecture that $r_d=2d-1$, which numerical experiments verify for $d\le100$.
+  We then provide an algorithm with $O\p{d^4}$ runtime for certifying upper bounds on the error with which tomography protocols can reconstruct an unknown quantum state.
+  Our algorithm motivates simple randomized tomography protocol, which we use to study the trade-off between the number of measurement axes and state reconstruction error [\red{note: this part is not yet complete}].
+  [\red{Possible things to look into: (1) a simplified tomography procedure for pure states, (2) adaptive tomography protocols.]}
 \end{abstract}
@@
 In particular, the protocols based on infinite-dimensional representations of a quantum spin require only the capability to measure spin projection onto different spatial axes, which is usually straightforward with a spin qudit.
-Nonetheless, these protocols obfuscate the requirements for performing full state tomography, provide no straightforward guarantees of state reconstruction fidelity, and extract only a vanishing faction of the information contained in measurement data.
+Nonetheless, these protocols obfuscate the requirements for performing full state tomography, provide no straightforward error bounds for reconstructed states, and extract only a vanishing faction of the information contained in measurement data.
 
@@
 In particular, this mapping allows us to bound the minimum number of measurement axes necessary to perform full tomography on a $d$-level spin qudit, $r_d$, by $2d-1\le r_d\le d^2$.
-We provide an example of an optimized tomography protocol in Section \ref{sec:protocol} [\red{I'll probably use modified versions of existing signal recovery algorithms rather than a new tomography protocol}], as well as numerical experiments to show that $r_d=2d-1$ for $d\le20$.
-
-We generally defer mathematical details and derivations to the appendices, keeping the main text focused on a high-level discussion of our results.
 
@@
 
-Moreover, any solution to the full recovery problem with a particular set of basis functions, such the spherical harmonics $Y_{\ell m}$, can be readily translated into a solution with a different set of basis functions, such as the adjoint rotation matrix elements $\D^\ell_{m,0}$ (see Appendix \ref{sec:basis_change}).
-This translation enables importing a host of existing signal recovery algorithms \cite{mcewen2011novel, mcewen2011sampling, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality} for the task of spin qudit tomography.
-Specifically, given a set of points $V\subset\Omega_2$ and coefficients $C_{\ell m}\p{\v v}$ that satisfy Eq.~\eqref{eq:full_recovery} with basis functions $Y_{\ell m}$, we can construct new coefficients $\tilde C_{\ell m}\p{\v v}$ that satisfy the same condition with basis functions $\D^\ell_{m,0}$ via
-\begin{align}
-  \tilde C_{\ell m} \equiv \sum_n G^\ell_{mn} C_{\ell n},
-\end{align}
-where
-\begin{align}
-  G^\ell_{mn} \equiv \f{2\ell+1}{4\pi}
-  \int_{\Omega_2} \d\Omega_2\p{\v v}
-  \D^\ell_{m,0}\p{\v v}^* Y_{\ell n}\p{\v v},
-\end{align}
-is an inner product of $\D^\ell_{m,0}$ and $Y_{\ell n}$.
-
-Finally, for any fixed degree $\ell$, finding solutions to the reduced recovery problem in Eq.~\eqref{eq:tomo_recovery} is equivalent to the recovery of a function in the degree-$\ell$ subspace of $L^2\p{\Omega_2}$, which is possible with $\abs{V}=2\ell+1$ samples (equal to the dimension of the degree-$\ell$ subspace) \cite{freeden2008spherical}.
+Moreover, for any fixed degree $\ell$, finding solutions to the reduced recovery problem in Eq.~\eqref{eq:tomo_recovery} is equivalent to the recovery of a function in the degree-$\ell$ subspace of $L^2\p{\Omega_2}$, which is possible with $\abs{V}=2\ell+1$ samples (equal to the dimension of the degree-$\ell$ subspace) \cite{freeden2008spherical}.
 In the case of spin qudit tomography, the degree $\ell$ takes a maximal value of $\ell_{\t{max}}\equiv d-1$, so state recovery requires at least as many measurement axes as there are transition operators with degree $\ell_{\t{max}}$, namely $2\ell_{\t{max}}+1=2d-1$.
@@
 Unfortunately, we have no general proof of this conjecture, which may require a deeper analysis of fixed-degree signal recovery on the sphere.
-Nonetheless, in the following section we show by example that $2d-1$ measurement axes suffice for all $d\le20$.
+Nonetheless, in the following section we show by example that $2d-1$ measurement axes suffice for all $d\le100$. [\red{modify sentence as appropriate after writing the next section}]
+
+In principle, any solution to the full recovery problem with a particular set of basis functions, such the spherical harmonics $Y_{\ell m}$, can be readily translated into a solution with a different set of basis functions, such as the adjoint rotation matrix elements $\D^\ell_{m,0}$ (see Appendix \ref{sec:basis_change}).
+This translation enables importing a host of existing signal recovery algorithms \cite{mcewen2011novel, mcewen2011sampling, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality} for the task of spin qudit tomography.
+In practice, spin qudits typically have only a modest dimension $d$, allowing for simpler and optimized tomography protocols that are practical despite worse scaling with $d$.
+We present one such protocol with $O\p{d^3}$ runtime in the following section.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Optimized tomography protocols}
+\section{Optimized tomography protocols and reconstruction error bounds}
 \label{sec:protocol}
@@
 \end{enumerate*}
-Whereas step \ref{tomo:measure} can involve a host of platform-dependent technical challenges, here we discuss the steps before and after measurement data collection.
+Whereas step \ref{tomo:measure} can involve a host of platform-dependent technical challenges, here we discuss the steps to take before and after measurement data collection.
 
-We begin by discussing the selection of measurement axes $V=\set{\v v}\subset\Omega_2$.
-In the absence of analytical tools to find ``optimal'' deterministic choice of axes, we compare three randomized selection methods that are motivated by past works on sparse signal recovery \cite{rauhut2011sparse, alem2012sparse}.
-Conceptually, the simplest method of randomly choosing axes would be to choose $\abs{V}$ points from the uniform distribution on the sphere.
-To select a point $\v v=\p{\alpha_{\v v},\beta_{\v v}}$ from the uniform distribution on the sphere, we first select a point $\p{a_{\v v},z_{\v v}}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and set $\alpha_{\v v}=2\pi a_{\v v}$ and $\beta_{\v v}=\arccos\p{1-2z_{\v v}}$.
-We will call this procedure the {\it uniform} sampling method.
+To this end, we begin by asking a question: what is a ``good'' choice of measurement axes?
+Intuitively, a good choice of axes should minimize the error with which one can reconstruct an unknown quantum state from associated measurement data.
+If we can quantify this idea, then we can try to optimize over different choices of measurement axes to find a set that (approximately) minimizes the error in reconstructed states.
 
-The work in Ref.~\cite{rauhut2011sparse} proposes an alternate sampling method for sparse signal recovery on the sphere, which consists of uniformly sampling angles $\p{\alpha_{\v v},\beta_{\v v}}\in[0,2\pi]\times[0,\pi]$.
-We will refer to this procedure as the {\it biased} sampling scheme, as it places a higher weight on points near the poles of the sphere.
-For general-purpose recovery of band-limited functions $f=\sum_{\ell,m} f_{\ell m} Y_{\ell m}$ with only a small number of non-zero spherical harmonic coefficients $f_{\ell m}$, Ref.~\cite{rauhut2011sparse} found that the biased sampling scheme was more likely to result in ``good'' points that allow for accurate signal recovery.
+A set of measurement axes $V=\set{\v v}$ nominally induces a set of projectors $\v\Pi_V=\set{\Pi_j^V}$ that will be measured in an experiment, where for shorthand we use a combined index $j\simeq\p{\v v_j,\mu_j}$ that specifies both a measurement axis $\v v_j$ and an outcome $\mu_j$, and for clarify we index objects by $V$ to emphasize their dependence on the choice of measurement axes.
+The projectors $\v\Pi_V$ can be used to construct the overlap (Gram) matrix
+\begin{align}
+  G_V = \sum_{i,j=1}^{\abs{V}\times d} \obk{\Pi_i^V|\Pi_j^V} \op{i}{j}.
+\end{align}
+Eigenvectors of $G_V$ define linearly independent linear combinations of the projectors in $\v\Pi_V$.
+If the projectors in $\v\Pi_V$ span the entire space of operators on a $d$-level spin qudit, which is a necessary and sufficient condition for $V$ to allow for full state tomography, then $G_V$ must have exactly $d^2$ nonzero eigenvalues.
+Indexing these eigenvalues $G^V_k$ by an integer $k\in\set{1,2,\cdots,d^2}$ and normalizing the corresponding eigenvectors $\v g^V_k\equiv(g^V_{k,1},g^V_{k,2},\cdots)$ as
+\begin{align}
+  \abs{\v g^V_k}^2 = \sum_{j=1}^{\abs{V}\times d} \abs{g^V_{kj}}
+  = \f1{G^V_k},
+\end{align}
+we can construct the orthonormal operators
+\begin{align}
+  Q^V_k \equiv \sum_j g^V_{kj} \Pi_j^V.
+\end{align}
+These operators allow us to expand any state $\rho$ of a $d$-level spin qudit in the form
+\begin{align}
+  \rho = \sum_{k=1}^{d^2} \rho_k^V Q_k^V,
+  &&
+  \rho_k^V \equiv \bk{Q^V_k}_\rho.
+\end{align}
+Given empirical estimates $E_j^V$ of the expectation values $\bk{\Pi_j^V}_\rho$, an empirical estimate $\tilde\rho$ of $\rho$ is then
+\begin{align}
+  \tilde\rho \equiv \sum_k \tilde\rho^V_k Q^V_k,
+\end{align}
+where
+\begin{align}
+  \tilde\rho^V_k \equiv \sum_j g^V_{kj} E_j^V
+  \approx \sum_j g^V_{kj} \bk{\Pi_j^V}_\rho
+  = \bk{Q^V_k}_\rho
+  = \rho^V_k.
+\end{align}
+The empirical estimate $\tilde\rho$ can be further improved by performing maximum-likelihood corrections, outlined in the ``algorithm for subproblem 1'' in Ref.~\cite{smolin2012efficient}, which essentially finds the closest state to $\tilde\rho$ with no negative eigenvalues.
 
-Finally, we consider the {\it spiral} sampling scheme laid out in Ref.~\cite{alem2012sparse}, which first generates a spiral of points that go from the north pole to the south, using this spiral to define a grid of points on the entire sphere, and then randomly choosing $\abs{V}$ points on this grid.
-
-[\red{todo: make figure visualizing these sampling schemes}]
+We can use the eigenvalues of the overlap matrix $G_V$ to place strict upper bounds on the distance between the empirical estimate $\tilde\rho$ and the true state $\rho$.
+If the estimates $E^V_j$ approximate $\bk{\Pi^V_j}_\rho$ to within an error $\epsilon$, i.e.
+\begin{align}
+  \abs{E^V_j-\bk{\Pi^V_j}_\rho} \le \epsilon,
+\end{align}
+then we can say that
+\begin{align}
+  \tilde\rho^V_k-\rho^V_k
+  = \sum_j g_{kj}^V \p{E^V_j-\bk{\Pi^V_j}_\rho}
+  \le \epsilon \sum_j g_{kj}^V \sigma_j,
+\end{align}
+where $\sigma_j\in\set{+1,-1}$ is the sign of $E^V_j-\bk{\Pi^V_j}_\rho$.
+Using the fact that the operators $Q^V_k$ are orthonormal, the 2-norm (Euclidean) distance $\norm{\tilde\rho-\rho}_2$ between $\tilde\rho$ and $\rho$ is therefore bounded as
+\begin{align}
+  \norm{\tilde\rho-\rho}_2^2
+  &= \obk{\tilde\rho-\rho|\tilde\rho-\rho}
+  = \sum_k \abs{\tilde\rho^V_k - \rho^V_k}^2 \nonumber \\
+  &\le \epsilon^2 \sum_k \abs{\sum_j g_{kj}^V \sigma_j}^2
+  = \epsilon^2 \sum_k \abs{\v g^V_k}^2
+  = \epsilon^2 \N_V^2,
+\end{align}
+where we used the fact that changing the signs of coefficients in $\v g^V_k$ does not affect its norm, and
+\begin{align}
+  \N_V^2 \equiv \sum_k \f1{G^V_k}.
+\end{align}
+Minimizing the norm $\N_V$ associated with a choice of measurement axes $V$ thus minimizes the error with which the empirical estimate $\tilde\rho$ approximates the true state $\rho$.
 
@@
 
-[\red{Think about maximizing the determinant of the coefficient matrix $C_{\ell m}\p{\v v}$?}]
+[\red{Mention past works on sparse signal recovery \cite{rauhut2011sparse, alem2012sparse}.}]
 
-[\red{Note: the text below is copy/pasted from a different set of notes.  I still need to adapt this text for this paper.}]
-
-Question: given a collection of projectors $\Pi\equiv\set{\Pi_j}$ and measurement outcomes $M\equiv\set{M_j}$, with $M_j$ an empirical estimate of $\bk{\Pi_j}_\rho=\tr\p{\rho\Pi_j}$, how do we determine whether the measurement data $M$ is sufficient to reconstruct $\rho$?
-Answer: reconstruction of $\rho$ is possible if $\Pi$ spans the entire space $\B_n$ of operators on the Hilbert space of an $n$-level spin.
-Given an orthonormal basis $\set{Q_\alpha}$ of self-adjoint operators spanning $\B_n$, we can expand
-\begin{align}
-  \rho = \sum_\alpha \rho_\alpha Q_\alpha,
-  &&
-  \rho_\alpha \equiv \bk{Q_\alpha}_\rho.
-\end{align}
-If $\Pi$ spans $\B_n$, then we can find a set of real numbers $c_{\alpha j}\in\RR$ for which
-\begin{align}
-  Q_\alpha = \sum_j c_{\alpha j} \Pi_j.
-\end{align}
-An empirical estimate $\tilde\rho$ of $\rho$ is then
-\begin{align}
-  \tilde\rho = \sum_\alpha \tilde\rho_\alpha Q_\alpha,
-\end{align}
-where\footnote{This estimate can be improved by defining $\tilde\rho$ as the least-squares fit to the set of linear equations $\tr\p{\tilde\rho\,\Pi_j} = M_j$, and performing maximum-likelihood corrections to $\tilde\rho$ \cite{smolin2012efficient}.}
-\begin{align}
-  \tilde\rho_\alpha \equiv \sum_j c_{\alpha j} M_j
-  \approx \sum_j c_{\alpha j} \bk{\Pi_j}_\rho
-  = \bk{Q_\alpha}_\rho
-  = \rho_\alpha.
-\end{align}
-As it turns out, numerical experiments with $n\le20$ suggest that in order to perfurm full tomography of $\rho$, it suffices to measure all projectors $\Pi_{\v v\mu}$ onto states of definite spin $\mu$ along $2n-1$ different axes $\v v$.
-The question remains: which axes should we choose?
-As a first pass, we can simply choose a set of $2n-1$ random axes, $V\equiv\set{\v v}$, by uniformly sampling points on the sphere.
-These random axes define a set of $\p{2n-1}\times n$ projectors $\Pi_V=\set{\Pi^V_j}$, where for shorthand we use a combined index $j=\p{\v v_j,\mu_j}$ that specifies both a measurement axis $\v v_j$ and an outcome $\mu_j$, and for clarity we index objects by $V$ to emphasize their dependence on the choice of measurement axes.
-The projectors $\Pi_V$ can be used to construct $n^2$ linearly independent operators by constructing a matrix $B_V$ with entries
-\begin{align}
-  \sp{B_V}_{ij} = \tr\p{\Pi^V_i \Pi^V_j}.
-\end{align}
-Indexing the (normalized) eigenvectors and the corresponding eigenvalues of $B_V$ by $\alpha$, respectively $\v x^V_\alpha$ and $m^V_\alpha$, for all $\alpha$ with eigenvalue $m^V_\alpha\ne0$ we can then construct the normalized and linearly independent operators
-\begin{align}
-  Q^V_\alpha \equiv \f1{\sqrt{m^V_\alpha}} \sum_j x^V_{\alpha j} \Pi^V_j.
-\end{align}
-To summarize: choosing a set $V$ of $2n-1$ random axes will determine a set of measurement operators $\Pi_V$ from which we can construct a complete basis $\set{Q^V_\alpha}$ of operators for $\B_n$.
-However, some choices of axes will be better than others: a good choice of axes $V$ will maximize the amount of information that the corresponding measurement data $M_V=\set{M^V_j}$ will contain about the expectation values $\bk{Q^V_\alpha}_\rho$ used to reconstruct $\rho$.
-The amount of information that $M_V$ contains about $\bk{Q^V_\alpha}_\rho$ is determined, in part, by the magnitude of the eigenvalue $m^V_\alpha$: if $m^V_\alpha$ is small, then $M_V$ will generally contain little information about $\bk{Q^V_\alpha}_\rho$, and vice versa\footnote{Think about the limit $m^V_\alpha\to0$, in which case the fact that $Q^V_\alpha$ is normalized implies that the operator $\sum_j x^V_{\alpha j} \Pi^V_j$ must have a vanishing norm.
-  A vanishing norm can only occur due to fine-tuned cancellations that make the expectation value of $\sum_j x^V_{\alpha j} \Pi^V_j$ difficult to estimate from noisy data on $\bk{\Pi^V_j}_\rho$.}.
-Rather than choosing $2n-1$ axes randomly, we can therefore choose a set of axes that minimizes some cost function $C\p{m_V}$ favoring large eigenvalues $m_V\equiv\set{m^V_\alpha}$.
-A reasonable cost function that treats all eigenvalues on equal footing, for example, is
-\begin{align}
-  C_{\t{inv}}\p{m_V} \equiv \sum_\alpha \f1{m^V_\alpha},
-\end{align}
-where the sum is implicitly performed over the $n^2$ largest eigenvalues of $B_V$ (as all other eigenvalues are guaranteed to be zero).
 Note that any information about $\rho$, obtained either from prior knowledge or preliminary measurement data, can be used to construct an even better choice of measurement axes, leading to the possibility of tailored or adaptive measurement protocols \cite{pereira2018adaptive} that are more efficient in terms of the number of measurements required to estimate $\rho$ to a fixed precision.

### 2020-10-23 sun_phases/spin_qudit_tomo.tex
clarify some points in the abstract

@@
   We conjecture that $r_d=2d-1$, which numerical experiments verify for $d\le100$.
-  We then provide an algorithm with $O\p{d^4}$ runtime for certifying upper bounds on the error with which tomography protocols can reconstruct an unknown quantum state.
+  We then provide an algorithm with $O\p{dr^3}$ runtime for certifying strict upper bounds on the error with which a tomography protocols with $r$ measurement axes can reconstruct an unknown quantum state.
   Our algorithm motivates simple randomized tomography protocol, which we use to study the trade-off between the number of measurement axes and state reconstruction error [\red{note: this part is not yet complete}].

### 2020-10-23 sun_phases/spin_qudit_tomo.tex
fixed mistake in the derivation of an error bound on a reconstructed state

@@
 
-\let\var\relax
-\DeclareMathOperator{\var}{var}
-
 \def\obra#1{\mathinner{({#1}|}}
@@
   By mapping this quantum tomography problem onto the classical problem of signal recovery on the sphere, we prove that full reconstruction of arbitrary mixed states is possible, and requires a minimum number of measurement axes, $r_d$, that is bounded by $2d-1\le r_d\le d^2$.
-  We conjecture that $r_d=2d-1$, which numerical experiments verify for $d\le100$.
+  We conjecture that $r_d=2d-1$, which we have verified for all $d\le100$ using numerical experiments.
   We then provide an algorithm with $O\p{dr^3}$ runtime for certifying strict upper bounds on the error with which a tomography protocols with $r$ measurement axes can reconstruct an unknown quantum state.
@@
 In practice, spin qudits typically have only a modest dimension $d$, allowing for simpler and optimized tomography protocols that are practical despite worse scaling with $d$.
-We present one such protocol with $O\p{d^3}$ runtime in the following section.
 
@@
 \begin{align}
-  \abs{\v g^V_k}^2 = \sum_{j=1}^{\abs{V}\times d} \abs{g^V_{kj}}
+  \abs{\v g^V_k}^2
+  = \sum_{j=1}^{\abs{V}\times d} \abs{g^V_{kj}}^2
   = \f1{G^V_k},
@@
 
-We can use the eigenvalues of the overlap matrix $G_V$ to place strict upper bounds on the distance between the empirical estimate $\tilde\rho$ and the true state $\rho$.
+The eigenvectors of the overlap matrix $G_V$ provide strict upper bounds on the distance between the empirical estimate $\tilde\rho$ and the true state $\rho$.
 If the estimates $E^V_j$ approximate $\bk{\Pi^V_j}_\rho$ to within an error $\epsilon$, i.e.
@@
 \end{align}
-then we can say that
+then the error in the estimate $\tilde\rho^V_k$ of $\rho^V_k$ is
 \begin{align}
-  \tilde\rho^V_k-\rho^V_k
-  = \sum_j g_{kj}^V \p{E^V_j-\bk{\Pi^V_j}_\rho}
-  \le \epsilon \sum_j g_{kj}^V \sigma_j,
+  \abs{\tilde\rho^V_k-\rho^V_k}
+  &= \abs{\sum_j g_{kj}^V \p{E^V_j-\bk{\Pi^V_j}_\rho}} \\
+  &\le \sum_j \abs{g_{kj}^V \p{E^V_j-\bk{\Pi^V_j}_\rho}} \\
+  &\le \epsilon \sum_j \abs{g_{kj}^V}
+  \equiv \epsilon^V_k.
 \end{align}
-where $\sigma_j\in\set{+1,-1}$ is the sign of $E^V_j-\bk{\Pi^V_j}_\rho$.
-Using the fact that the operators $Q^V_k$ are orthonormal, the 2-norm (Euclidean) distance $\norm{\tilde\rho-\rho}_2$ between $\tilde\rho$ and $\rho$ is therefore bounded as
+Using the fact that the operators $Q^V_k$ are orthonormal, we can bound the 2-norm (Euclidean) distance $\norm{\tilde\rho-\rho}_2$ between $\tilde\rho$ and $\rho$ as
 \begin{align}
   \norm{\tilde\rho-\rho}_2^2
-  &= \obk{\tilde\rho-\rho|\tilde\rho-\rho}
-  = \sum_k \abs{\tilde\rho^V_k - \rho^V_k}^2 \nonumber \\
-  &\le \epsilon^2 \sum_k \abs{\sum_j g_{kj}^V \sigma_j}^2
-  = \epsilon^2 \sum_k \abs{\v g^V_k}^2
-  = \epsilon^2 \N_V^2,
+  = \sum_k \abs{\tilde\rho^V_k - \rho^V_k}^2
+  \le \sum_k \abs{\epsilon_k^V}^2
+  = \abs{\v\epsilon_V}^2,
 \end{align}
-where we used the fact that changing the signs of coefficients in $\v g^V_k$ does not affect its norm, and
-\begin{align}
-  \N_V^2 \equiv \sum_k \f1{G^V_k}.
-\end{align}
-Minimizing the norm $\N_V$ associated with a choice of measurement axes $V$ thus minimizes the error with which the empirical estimate $\tilde\rho$ approximates the true state $\rho$.
+where the vector $\v\epsilon_V \equiv \p{\epsilon^V_1,\epsilon^V_2,\cdots,\epsilon^V_{d^2}}$ depends only on the approximation error $\epsilon$ and choice of measurement axes $V$.
+To better understand the behavior of this bound,
 

### 2020-10-23 sun_phases/spin_qudit_tomo.tex
cleaned up and interpreted error bounds on reconstruction error

@@
   We conjecture that $r_d=2d-1$, which we have verified for all $d\le100$ using numerical experiments.
-  We then provide an algorithm with $O\p{dr^3}$ runtime for certifying strict upper bounds on the error with which a tomography protocols with $r$ measurement axes can reconstruct an unknown quantum state.
+  We then provide an algorithm with $O\p{dr^3}$ serial runtime for certifying strict upper bounds on the error with which a tomography protocols with $r$ measurement axes can reconstruct an unknown quantum state.
   Our algorithm motivates simple randomized tomography protocol, which we use to study the trade-off between the number of measurement axes and state reconstruction error [\red{note: this part is not yet complete}].
@@
 If the projectors in $\v\Pi_V$ span the entire space of operators on a $d$-level spin qudit, which is a necessary and sufficient condition for $V$ to allow for full state tomography, then $G_V$ must have exactly $d^2$ nonzero eigenvalues.
-Indexing these eigenvalues $G^V_k$ by an integer $k\in\set{1,2,\cdots,d^2}$ and normalizing the corresponding eigenvectors $\v g^V_k\equiv(g^V_{k,1},g^V_{k,2},\cdots)$ as
-\begin{align}
-  \abs{\v g^V_k}^2
-  = \sum_{j=1}^{\abs{V}\times d} \abs{g^V_{kj}}^2
-  = \f1{G^V_k},
-\end{align}
+Indexing these eigenvalues $G^V_k$ and the corresponding (normalized) eigenvectors $\v x^V_k\equiv(x^V_{k,1},x^V_{k,2},\cdots)$ by an integer $k\in\set{1,2,\cdots,d^2}$,
 we can construct the orthonormal operators
 \begin{align}
-  Q^V_k \equiv \sum_j g^V_{kj} \Pi_j^V.
+  Q^V_k \equiv \sum_j g^V_{kj} \Pi_j^V,
+  &&
+  g^V_{kj} \equiv \f{x^V_{kj}}{\sqrt{G^V_k}}.
 \end{align}
@@
   &\le \sum_j \abs{g_{kj}^V \p{E^V_j-\bk{\Pi^V_j}_\rho}} \\
-  &\le \epsilon \sum_j \abs{g_{kj}^V}
-  \equiv \epsilon^V_k.
+  &\le \epsilon \norm{\v g^V_k}_1,
 \end{align}
+where $\norm{\v g}_1=\sum_j\abs{g_j}$ is the 1-norm of the vector $\v g$.
 Using the fact that the operators $Q^V_k$ are orthonormal, we can bound the 2-norm (Euclidean) distance $\norm{\tilde\rho-\rho}_2$ between $\tilde\rho$ and $\rho$ as
@@
   = \sum_k \abs{\tilde\rho^V_k - \rho^V_k}^2
-  \le \sum_k \abs{\epsilon_k^V}^2
-  = \abs{\v\epsilon_V}^2,
+  \le \epsilon^2 \N_V^2,
 \end{align}
-where the vector $\v\epsilon_V \equiv \p{\epsilon^V_1,\epsilon^V_2,\cdots,\epsilon^V_{d^2}}$ depends only on the approximation error $\epsilon$ and choice of measurement axes $V$.
-To better understand the behavior of this bound,
+where
+\begin{align}
+  \N_V^2 \equiv \min_U \sum_k \norm{U\v g^V_k}_1^2
+  = \min_U \sum_k \f{\norm{U\v x^V_k}_1^2}{G^V_k}
+  \label{eq:norm}
+\end{align}
+is a norm uniquely determined by the set of measurement axes $V$, with minimization performed over unitary transformations $U$ of all $\v g^V_k$ and $\v x^V_k$ to account for the fact that their 1-norms are generally basis-dependent.
+
+To better understand this norm, which induces a state-agnostic upper bound on reconstruction error, we first observe that by a change of operator basis, measuring the projectors $\set{\Pi_{\v v\mu}}$ is equivalent to measuring the transition operators $\set{T_{\v v\ell,0}}$.
+Orthogonality of transition operators with different degrees $\ell$ then implies that the overlap matrix $G_V$ has a block-diagonal structure in the transition operator basis, with $d$ blocks of size $\abs{V}\times\abs{V}$.
+Eigenvectors of $G_V$ in this basis can therefore have at most $\abs{V}$ nonzero components, which implies that
+\begin{align}
+  \min_U \norm{U\v x^V_k}_1^2
+  \le \abs{V}\, \norm{\v x^V_k}_2^2
+  = \abs{V},
+\end{align}
+so
+\begin{align}
+  \N_V^2 \le \abs{V} \sum_k \f1{G^V_k}.
+\end{align}
+Note that this sum, as well as that in Eq.~\eqref{eq:norm}, is performed over the $d^2$ largest eigenvalues of $G_V$, and therefore may be divergent if $G_V$ has fewer than $d^2$ nonzero eigenvalues.
+Altogether, diagonalizing each of $d$ blocks of $G_V$ takes a time that is $O(\abs{V}^3)$, so an upper bound for $\N_V$ can be computed in $O(d\abs{V}^3)$ serial or $O(\abs{V}^3)$ parallel runtime.
 

### 2020-10-23 sun_phases/spin_qudit_tomo.tex
clarify reason for minimization over unitaries in error bound

@@
   G_V = \sum_{i,j=1}^{\abs{V}\times d} \obk{\Pi_i^V|\Pi_j^V} \op{i}{j}.
+  \label{eq:overlap_matrix}
 \end{align}
@@
   \le \epsilon^2 \N_V^2,
+  \label{eq:bound}
 \end{align}
@@
 is a norm uniquely determined by the set of measurement axes $V$, with minimization performed over unitary transformations $U$ of all $\v g^V_k$ and $\v x^V_k$ to account for the fact that their 1-norms are generally basis-dependent.
+Equivalently, this minimization accounts for the fact that we can perform the analysis between Eqs.~\eqref{eq:overlap_matrix} and \eqref{eq:bound} identically with any $G_V\to U G_V U^\dag$, where $U$ is an arbitrary unitary that mixes the projectors in $\v\Pi_V$.
 
-To better understand this norm, which induces a state-agnostic upper bound on reconstruction error, we first observe that by a change of operator basis, measuring the projectors $\set{\Pi_{\v v\mu}}$ is equivalent to measuring the transition operators $\set{T_{\v v\ell,0}}$.
-Orthogonality of transition operators with different degrees $\ell$ then implies that the overlap matrix $G_V$ has a block-diagonal structure in the transition operator basis, with $d$ blocks of size $\abs{V}\times\abs{V}$.
-Eigenvectors of $G_V$ in this basis can therefore have at most $\abs{V}$ nonzero components, which implies that
+To better understand the bound in Eq.~\eqref{eq:bound}, we observe that by a change of operator basis, or equivalently a particular choice of unitary $U=U_0$, measuring the projectors $\set{\Pi_{\v v\mu}}$ is equivalent to measuring the transition operators $\set{T_{\v v\ell,0}}$.
+Orthogonality of transition operators with different degrees $\ell$ then implies that the overlap matrix $U_0 G_V U_0^\dag$ has a block-diagonal structure, with $d$ blocks of size $\abs{V}\times\abs{V}$.
+Eigenvectors of $G_V$ in the transition operator basis can therefore have at most $\abs{V}$ nonzero components, which implies that
 \begin{align}
   \min_U \norm{U\v x^V_k}_1^2
+  \le \norm{U_0\v x^V_k}_1^2
   \le \abs{V}\, \norm{\v x^V_k}_2^2

### 2020-10-23 sun_phases/spin_qudit_tomo.tex
added discussion of an explicit tomography protocol

@@
 \item select a set of measurement axes,
-  \label{tomo:select}
 \item collect measurement data on spin projection onto these axes, and then
@@
 \item process the collected data to reconstruct the state of the spin qudit.
-  \label{tomo:process}
 \end{enumerate*}
@@
   \tilde\rho \equiv \sum_k \tilde\rho^V_k Q^V_k,
+  \label{eq:reconstructed_state}
 \end{align}
@@
 \end{align}
-The empirical estimate $\tilde\rho$ can be further improved by performing maximum-likelihood corrections, outlined in the ``algorithm for subproblem 1'' in Ref.~\cite{smolin2012efficient}, which essentially finds the closest state to $\tilde\rho$ with no negative eigenvalues.
-
 The eigenvectors of the overlap matrix $G_V$ provide strict upper bounds on the distance between the empirical estimate $\tilde\rho$ and the true state $\rho$.
@@
 
-To better understand the bound in Eq.~\eqref{eq:bound}, we observe that by a change of operator basis, or equivalently a particular choice of unitary $U=U_0$, measuring the projectors $\set{\Pi_{\v v\mu}}$ is equivalent to measuring the transition operators $\set{T_{\v v\ell,0}}$.
-Orthogonality of transition operators with different degrees $\ell$ then implies that the overlap matrix $U_0 G_V U_0^\dag$ has a block-diagonal structure, with $d$ blocks of size $\abs{V}\times\abs{V}$.
-Eigenvectors of $G_V$ in the transition operator basis can therefore have at most $\abs{V}$ nonzero components, which implies that
+To better understand the bound in Eq.~\eqref{eq:bound}, we observe that by a change of operator basis, or equivalently a particular choice of unitary $U=U_T$, measuring the projectors $\set{\Pi_{\v v\mu}}$ is equivalent to measuring the transition operators $\set{T_{\v v\ell,0}}$.
+Orthogonality of transition operators with different degrees $\ell$ then implies that the overlap matrix
+\begin{align}
+  G_V^{\t{trn}} \equiv U_T G_V U_T^\dag
+  = \sum_{\ell,\v v,\v v'} \obk{T_{\v v\ell,0}|T_{\v v'\ell,0}}
+  \op{\v v\ell}{\v v'\ell}
+\end{align}
+has a block-diagonal structure, with $d$ blocks of size $\abs{V}\times\abs{V}$.
+Eigenvectors of $G_V$ can therefore have at most $\abs{V}$ nonzero components in the transition operator basis, which implies that
 \begin{align}
   \min_U \norm{U\v x^V_k}_1^2
-  \le \norm{U_0\v x^V_k}_1^2
+  \le \norm{U_T\v x^V_k}_1^2
   \le \abs{V}\, \norm{\v x^V_k}_2^2
@@
 
-\vspace{3cm}
+The bound in Eq.~\eqref{eq:bound} motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere, and variationally optimize the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the cost function
+\begin{align}
+  \tilde\N_V^2 \equiv \sum_k \f{\norm{U_T\v x^V_k}_1^2}{G^V_k}
+  \ge \N_V^2,
+\end{align}
+where $U_T\v x^V_k$ are (normalized) eigenvectors of the overlap matrix $G_V$ in the transition operator basis, obtained by diagonalizing $G_V^{\t{trn}}$.
+If $\abs{V}$ is too large for variational optimization to be practical, one can instead generate many random sets of measurement axes, and then choose the set that minimizes $\tilde\N_V^2$.
+Upon choosing a set of measurement axes $V$, experimentally characterize the expectation values $\bk{\Pi^V_j}_\rho$.
+If these measurements can approximate the expectation values $\bk{\Pi^V_j}_\rho$ to within a known error $\epsilon$, then the empirical estimate $\tilde\rho$ defined in Eq.~\eqref{eq:reconstructed_state} is an approximation of the true state $\rho$ to within a 2-norm distance of $\epsilon\N_V$.
+If the empirical estimate $\tilde\rho$ has negative eigenvalues, its 2-norm distance from $\rho$ can be further reduced via maximum-likelihood corrections \cite{smolin2012efficient}.
+In terms of classical computing costs before and after the estimation of expectation values $\bk{\Pi^V_j}_\rho$, this tomography protocol is bottlenecked by the need to diagonalize the overlap matrix $G_V^{\t{trn}}$, which has $d$ blocks of size $\abs{V}\times\abs{V}$.
+Altogether, this protocol has an $O(d\abs{V}^3)$ serial or $O(\abs{V}^3)$ parallel runtime.
 
-[\red{Mention past works on sparse signal recovery \cite{rauhut2011sparse, alem2012sparse}.}]
+Note that any information about $\rho$, obtained either from prior knowledge or preliminary measurement data, can be used to construct tailored or adaptive measurement protocols \cite{pereira2018adaptive} that are more efficient in terms of the number of measurements required to estimate $\rho$ to a fixed precision.
+We leave the development of tailored and adaptive measurement protocols to future work.
 
-Note that any information about $\rho$, obtained either from prior knowledge or preliminary measurement data, can be used to construct an even better choice of measurement axes, leading to the possibility of tailored or adaptive measurement protocols \cite{pereira2018adaptive} that are more efficient in terms of the number of measurements required to estimate $\rho$ to a fixed precision.
-We leave the details of tailored and adaptive measurement protocols to future work.
+[\red{Mention past works on sparse signal recovery \cite{rauhut2011sparse, alem2012sparse}?}]
 

### 2020-10-23 sun_phases/spin_qudit_tomo.tex
added paragragh about sparse signal recovery

@@
   We conjecture that $r_d=2d-1$, which we have verified for all $d\le100$ using numerical experiments.
-  We then provide an algorithm with $O\p{dr^3}$ serial runtime for certifying strict upper bounds on the error with which a tomography protocols with $r$ measurement axes can reconstruct an unknown quantum state.
-  Our algorithm motivates simple randomized tomography protocol, which we use to study the trade-off between the number of measurement axes and state reconstruction error [\red{note: this part is not yet complete}].
-  [\red{Possible things to look into: (1) a simplified tomography procedure for pure states, (2) adaptive tomography protocols.]}
+  We then provide an algorithm with $O\p{r^3}$ runtime for certifying strict upper bounds on the error with which measurements of spin projection along $r$ measurement axes can be used to reconstruct an unknown quantum state.
+  Our algorithm motivates a simple randomized tomography protocol, which we use to study the trade-off between the number of measurement axes and state reconstruction error [\red{note: this part is not yet complete}].
+  [\red{todo: maybe look into a simplified tomography procedure for pure states?]}
 \end{abstract}
@@
 is a norm uniquely determined by the set of measurement axes $V$, with minimization performed over unitary transformations $U$ of all $\v g^V_k$ and $\v x^V_k$ to account for the fact that their 1-norms are generally basis-dependent.
-Equivalently, this minimization accounts for the fact that we can perform the analysis between Eqs.~\eqref{eq:overlap_matrix} and \eqref{eq:bound} identically with any $G_V\to U G_V U^\dag$, where $U$ is an arbitrary unitary that mixes the projectors in $\v\Pi_V$.
+Seen otherwise, this minimization accounts for the fact that we can perform the analysis between Eqs.~\eqref{eq:overlap_matrix} and \eqref{eq:bound} identically with any $G_V\to U G_V U^\dag$, where $U$ is an arbitrary unitary that mixes the projectors in $\v\Pi_V$.
 
@@
 
-The bound in Eq.~\eqref{eq:bound} motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere, and variationally optimize the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the cost function
+The bound in Eq.~\eqref{eq:bound} motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and variationally optimize the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the cost function
 \begin{align}
@@
 
-[\red{Mention past works on sparse signal recovery \cite{rauhut2011sparse, alem2012sparse}?}]
+As a last point, we mention that the cost function $\tilde\N_V^2$ can also evaluate the quality of measurement axes chosen using existing methods for sparse signal recovery on the sphere \cite{rauhut2011sparse, alem2012sparse}.
+In practice, we find that uniform random sampling outperforms these methods, in the sense that it generates low-cost measurement axes with higher probability.
+Two observations may explain the poor performance of the sparse signal recovery methods: first, that spin qudit tomography essentially boils down to recovering sparse signals with a particular harmonic structure, namely with a fixed degree $\ell$.
+The existing sparse signal recovery algorithms do not assume or account for any harmonic structure, and may therefore perform worse than expected on this particular subset of sparse signals.
+Second, spin qudit tomography essentially requires reconstructing signals that are band-limited at degree $d$, but have $O(d)$ non-zero harmonic coefficients (out of $d^2$ coefficients in total).
+The existing sparse signal recovery methods, on the other hand, may only work for signals with sparsity much smaller than $d$.
+In any case, the fact that uniform sampling outperforms these methods only simplifies our randomized tomography protocol.
 

### 2020-10-23 sun_phases/spin_qudit_tomo.tex
minor wording changes

@@
 \end{enumerate*}
-Whereas step \ref{tomo:measure} can involve a host of platform-dependent technical challenges, here we discuss the steps to take before and after measurement data collection.
+Whereas step \ref{tomo:measure} can involve a host of platform-dependent technical challenges, here we discuss the steps to take before and after collecting measurement data.
 
@@
 \end{align}
-The eigenvectors of the overlap matrix $G_V$ provide strict upper bounds on the distance between the empirical estimate $\tilde\rho$ and the true state $\rho$.
+The eigenvalues and eigenvectors of the overlap matrix $G_V$ provide strict upper bounds on the distance between the empirical estimate $\tilde\rho$ and the true state $\rho$.
 If the estimates $E^V_j$ approximate $\bk{\Pi^V_j}_\rho$ to within an error $\epsilon$, i.e.
@@
 The existing sparse signal recovery algorithms do not assume or account for any harmonic structure, and may therefore perform worse than expected on this particular subset of sparse signals.
-Second, spin qudit tomography essentially requires reconstructing signals that are band-limited at degree $d$, but have $O(d)$ non-zero harmonic coefficients (out of $d^2$ coefficients in total).
-The existing sparse signal recovery methods, on the other hand, may only work for signals with sparsity much smaller than $d$.
+Second, spin qudit tomography essentially requires recovering signals that are band-limited at degree $d$, but have $O(d)$ non-zero harmonic coefficients (out of $d^2$ coefficients in total).
+The existing sparse signal recovery methods, on the other hand, may only perform well for signals with sparsity much smaller than $d$.
 In any case, the fact that uniform sampling outperforms these methods only simplifies our randomized tomography protocol.

### 2020-10-23 sun_phases/spin_qudit_tomo.tex
added note concerning a tightened error bound

@@
 \newcommand{\B}{\mathcal{B}}
+\newcommand{\M}{\mathcal{M}}
 \newcommand{\N}{\mathcal{N}}
@@
 is a norm uniquely determined by the set of measurement axes $V$, with minimization performed over unitary transformations $U$ of all $\v g^V_k$ and $\v x^V_k$ to account for the fact that their 1-norms are generally basis-dependent.
-Seen otherwise, this minimization accounts for the fact that we can perform the analysis between Eqs.~\eqref{eq:overlap_matrix} and \eqref{eq:bound} identically with any $G_V\to U G_V U^\dag$, where $U$ is an arbitrary unitary that mixes the projectors in $\v\Pi_V$.
+Seen otherwise, this minimization accounts for the fact that we can repeat the analysis between Eqs.~\eqref{eq:overlap_matrix} and \eqref{eq:bound} with any $G_V\to U G_V U^\dag$, where $U$ is an arbitrary unitary that mixes the projectors in $\v\Pi_V$.
 
-To better understand the bound in Eq.~\eqref{eq:bound}, we observe that by a change of operator basis, or equivalently a particular choice of unitary $U=U_T$, measuring the projectors $\set{\Pi_{\v v\mu}}$ is equivalent to measuring the transition operators $\set{T_{\v v\ell,0}}$.
+To better understand the bound in Eq.~\eqref{eq:bound}, we observe that by a change of operator basis, or equivalently a particular choice of unitary $U_T$, measuring the projectors $\set{\Pi_{\v v\mu}}$ is equivalent to measuring the transition operators $\set{T_{\v v\ell,0}}$.
 Orthogonality of transition operators with different degrees $\ell$ then implies that the overlap matrix
@@
 \begin{align}
-  \min_U \norm{U\v x^V_k}_1^2
-  \le \norm{U_T\v x^V_k}_1^2
-  \le \abs{V}\, \norm{\v x^V_k}_2^2
-  = \abs{V},
+  \N_V \le \sum_k \f{\norm{U_T\v x^V_k}_1^2}{G^V_k}
+  \le \sum_k \f{\abs{V}\,\norm{U_T\v x^V_k}_2^2}{G^V_k}
+  = \sum_k \f{\abs{V}}{G^V_k}.
 \end{align}
-so
-\begin{align}
-  \N_V^2 \le \abs{V} \sum_k \f1{G^V_k}.
-\end{align}
-Note that this sum, as well as that in Eq.~\eqref{eq:norm}, is performed over the $d^2$ largest eigenvalues of $G_V$, and therefore may be divergent if $G_V$ has fewer than $d^2$ nonzero eigenvalues.
-Altogether, diagonalizing each of $d$ blocks of $G_V$ takes a time that is $O(\abs{V}^3)$, so an upper bound for $\N_V$ can be computed in $O(d\abs{V}^3)$ serial or $O(\abs{V}^3)$ parallel runtime.
+Note that these sums are performed over the $d^2$ largest eigenvalues of $G_V$ (or $G_V^{\t{trn}}$), and therefore may be divergent if $G_V$ has fewer than $d^2$ nonzero eigenvalues.
+Altogether, diagonalizing each of $d$ blocks of $G_V^{\t{trn}}$ takes a time that is $O(\abs{V}^3)$, so an upper bound for $\N_V$ can be computed in $O(d\abs{V}^3)$ serial or $O(\abs{V}^3)$ parallel runtime.
+[\red{Note: rather than an absolute error $\epsilon$ on experimental measurements, we can assume that $E_j$ is equal to $\bk{\Pi_j}_\rho$ up to Gaussian noise with variance $\epsilon^2$.
+In that case, rather than a strict error bound we can make statements about the expected (squared) distance between $\tilde\rho$ and $\rho$, which turns out to be simply $\epsilon^2\sum_k1/G^V_k$.
+In retrospect, I think that this is a much nicer result, but I haven't written it up yet, which will require me to restructure this section somewhat.}]
 
@@
 \begin{align}
-  \tilde\N_V^2 \equiv \sum_k \f{\norm{U_T\v x^V_k}_1^2}{G^V_k}
+  \tilde\N_V^2 \equiv \sum_k \f{\norm{\v y^V_k}_1^2}{G^V_k}
   \ge \N_V^2,
 \end{align}
-where $U_T\v x^V_k$ are (normalized) eigenvectors of the overlap matrix $G_V$ in the transition operator basis, obtained by diagonalizing $G_V^{\t{trn}}$.
+where $\v y^V_k$ are the (normalized) eigenvectors of the overlap matrix $G_V^{\t{trn}}$ in the transition operator basis, with corresponding eigenvalues $G^V_k$.
 If $\abs{V}$ is too large for variational optimization to be practical, one can instead generate many random sets of measurement axes, and then choose the set that minimizes $\tilde\N_V^2$.

### 2020-10-23 sun_phases/spin_qudit_tomo.tex
last minor changes before overhauling reconstruction error bounds

@@
 \renewcommand{\i}{\mathrm{i}\mkern1mu} % imaginary unit
-
-\usepackage{dsfont} % for identity operator
-\newcommand{\1}{\mathds{1}}
-
-\usepackage{mathtools} % for \coloneqq
 
@@
   By mapping this quantum tomography problem onto the classical problem of signal recovery on the sphere, we prove that full reconstruction of arbitrary mixed states is possible, and requires a minimum number of measurement axes, $r_d$, that is bounded by $2d-1\le r_d\le d^2$.
-  We conjecture that $r_d=2d-1$, which we have verified for all $d\le100$ using numerical experiments.
+  We conjecture that $r_d=2d-1$, which we have verified for all $d\le120$ using numerical experiments.
   We then provide an algorithm with $O\p{r^3}$ runtime for certifying strict upper bounds on the error with which measurements of spin projection along $r$ measurement axes can be used to reconstruct an unknown quantum state.
@@
 Unfortunately, we have no general proof of this conjecture, which may require a deeper analysis of fixed-degree signal recovery on the sphere.
-Nonetheless, in the following section we show by example that $2d-1$ measurement axes suffice for all $d\le100$. [\red{modify sentence as appropriate after writing the next section}]
+Nonetheless, we have used the tomography protocol in the next section to numerically verify that $2d-1$ measurement axes suffice for all $d\le120$.
 
@@
 Seen otherwise, this minimization accounts for the fact that we can repeat the analysis between Eqs.~\eqref{eq:overlap_matrix} and \eqref{eq:bound} with any $G_V\to U G_V U^\dag$, where $U$ is an arbitrary unitary that mixes the projectors in $\v\Pi_V$.
+[\red{Actually, I don't think this minimization is quite correct as stated, but it won't matter after some pending changes to this section.
+Rather than an absolute error $\epsilon$, we can assume that $E_j$ approximates $\bk{\Pi_j}_\rho$ up to Gaussian noise with variance $\epsilon^2$.
+In that case, I think the mean (squared) distance between $\tilde\rho$ and $\rho$ is $\epsilon^2\sum_k1/G^V_k$, which in retrospect is a much nicer result.
+I am still working to write this up, which will involve some changes to this section.}]
 
@@
 Altogether, diagonalizing each of $d$ blocks of $G_V^{\t{trn}}$ takes a time that is $O(\abs{V}^3)$, so an upper bound for $\N_V$ can be computed in $O(d\abs{V}^3)$ serial or $O(\abs{V}^3)$ parallel runtime.
-[\red{Note: rather than an absolute error $\epsilon$ on experimental measurements, we can assume that $E_j$ is equal to $\bk{\Pi_j}_\rho$ up to Gaussian noise with variance $\epsilon^2$.
-In that case, rather than a strict error bound we can make statements about the expected (squared) distance between $\tilde\rho$ and $\rho$, which turns out to be simply $\epsilon^2\sum_k1/G^V_k$.
-In retrospect, I think that this is a much nicer result, but I haven't written it up yet, which will require me to restructure this section somewhat.}]
 
@@
 If the empirical estimate $\tilde\rho$ has negative eigenvalues, its 2-norm distance from $\rho$ can be further reduced via maximum-likelihood corrections \cite{smolin2012efficient}.
-In terms of classical computing costs before and after the estimation of expectation values $\bk{\Pi^V_j}_\rho$, this tomography protocol is bottlenecked by the need to diagonalize the overlap matrix $G_V^{\t{trn}}$, which has $d$ blocks of size $\abs{V}\times\abs{V}$.
-Altogether, this protocol has an $O(d\abs{V}^3)$ serial or $O(\abs{V}^3)$ parallel runtime.
+In terms of classical computing costs before and after estimating the expectation values $\bk{\Pi^V_j}_\rho$, this tomography protocol is bottlenecked by the need to diagonalize the overlap matrix $G_V^{\t{trn}}$, which has $d$ blocks of size $\abs{V}\times\abs{V}$.
+Altogether, this protocol therefore has an $O(d\abs{V}^3)$ serial or $O(\abs{V}^3)$ parallel runtime.
 
@@
 As a last point, we mention that the cost function $\tilde\N_V^2$ can also evaluate the quality of measurement axes chosen using existing methods for sparse signal recovery on the sphere \cite{rauhut2011sparse, alem2012sparse}.
-In practice, we find that uniform random sampling outperforms these methods, in the sense that it generates low-cost measurement axes with higher probability.
-Two observations may explain the poor performance of the sparse signal recovery methods: first, that spin qudit tomography essentially boils down to recovering sparse signals with a particular harmonic structure, namely with a fixed degree $\ell$.
+In practice, we find that uniform random sampling outperforms these methods, in the sense that it generates low-cost sets of measurement axes with higher probability.
+Two observations may explain the poor performance of the sparse signal recovery methods.
+First, spin qudit tomography essentially boils down to recovering sparse signals with a particular harmonic structure, namely with a fixed degree $\ell$.
 The existing sparse signal recovery algorithms do not assume or account for any harmonic structure, and may therefore perform worse than expected on this particular subset of sparse signals.
-Second, spin qudit tomography essentially requires recovering signals that are band-limited at degree $d$, but have $O(d)$ non-zero harmonic coefficients (out of $d^2$ coefficients in total).
+Second, spin qudit tomography requires recovering signals that are band-limited at degree $d$, but have $O(d)$ non-zero harmonic coefficients (out of $d^2$ coefficients in total).
 The existing sparse signal recovery methods, on the other hand, may only perform well for signals with sparsity much smaller than $d$.
 In any case, the fact that uniform sampling outperforms these methods only simplifies our randomized tomography protocol.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{How many measurement axes?}
+\label{sec:axes}
 

### 2020-10-24 sun_phases/spin_qudit_tomo.tex
overhaul and simplify reconstruction error bounds

@@
 
+% double angle brackets
+\newcommand{\bbk}[1]{\langle\!\langle #1 \rangle\!\rangle}
+\newcommand{\Bbk}[1]
+{\left\langle\!\!\left\langle #1 \right\rangle\!\!\right\rangle}
+
+\usepackage{dsfont} % for identity operator
+\newcommand{\1}{\mathds{1}}
+
 \newcommand{\up}{\uparrow}
@@
 \newcommand{\B}{\mathcal{B}}
-\newcommand{\M}{\mathcal{M}}
 \newcommand{\N}{\mathcal{N}}
@@
 
-A set of measurement axes $V=\set{\v v}$ nominally induces a set of projectors $\v\Pi_V=\set{\Pi_j^V}$ that will be measured in an experiment, where for shorthand we use a combined index $j\simeq\p{\v v_j,\mu_j}$ that specifies both a measurement axis $\v v_j$ and an outcome $\mu_j$, and for clarify we index objects by $V$ to emphasize their dependence on the choice of measurement axes.
-The projectors $\v\Pi_V$ can be used to construct the overlap (Gram) matrix
+A set of measurement axes $V=\set{\v v}$ nominally induces a set of projectors $\set{\Pi_{\v v\mu}^V}$ that will be measured in an experiment, where for clarify we index objects by $V$ to emphasize their dependence on the choice of measurement axes.
+By a simple change of basis, measuring these projectors is equivalent to measuring the set of transition operators $\v T_V\equiv\set{T_{\v v\ell,0}}$.
+The operators in $\v T_V$ can be used to construct the overlap (Gram) matrix
 \begin{align}
-  G_V = \sum_{i,j=1}^{\abs{V}\times d} \obk{\Pi_i^V|\Pi_j^V} \op{i}{j}.
+  G_V \equiv \sum_{\v v,\v v',\ell}
+  \obk{T_{\v v\ell,0}|T_{\v v'\ell,0}} \op{\v v\ell}{\v v'\ell},
   \label{eq:overlap_matrix}
 \end{align}
-Eigenvectors of $G_V$ define linearly independent linear combinations of the projectors in $\v\Pi_V$.
-If the projectors in $\v\Pi_V$ span the entire space of operators on a $d$-level spin qudit, which is a necessary and sufficient condition for $V$ to allow for full state tomography, then $G_V$ must have exactly $d^2$ nonzero eigenvalues.
-Indexing these eigenvalues $G^V_k$ and the corresponding (normalized) eigenvectors $\v x^V_k\equiv(x^V_{k,1},x^V_{k,2},\cdots)$ by an integer $k\in\set{1,2,\cdots,d^2}$,
-we can construct the orthonormal operators
+where we neglect cross terms with different degrees $\ell$ because the corresponding inner products $\obk{T_{\v v\ell,0}|T_{\v v'\ell',0}}=0$ for all $\ell\ne\ell'$.
+Eigenvectors of $G_V$ define linearly independent linear combinations of the operators in $\v T_V$.
+If the operators in $\v T_V$ span the entire space of operators on a $d$-level spin qudit, which is a necessary and sufficient condition for $V$ to allow for full state tomography, then $G_V$ must have exactly $d^2$ nonzero eigenvalues.
+Indexing these eigenvalues $G^V_k$ and the corresponding (normalized) eigenvectors $\v x^V_k\equiv(x^V_{k,1},x^V_{k,2},\cdots)$ by an integer $k\in\set{1,2,\cdots,\abs{V}d}$, we can construct the orthonormal operators
 \begin{align}
-  Q^V_k \equiv \sum_j g^V_{kj} \Pi_j^V,
+  Q^V_k \equiv \sum_j g^V_{kj} T_j,
   &&
-  g^V_{kj} \equiv \f{x^V_{kj}}{\sqrt{G^V_k}}.
+  g^V_{kj} \equiv \f{x^V_{kj}}{\sqrt{G^V_k}},
 \end{align}
+where for shorthand we use a combined index $j\equiv\p{\v v_j,\ell_j}$ to specify both a measurement axis $\v v_j$ and a degree $\ell_j$, which identify the transition operator $T_j\equiv T_{\v v_j\ell_j,0}$.
 These operators allow us to expand any state $\rho$ of a $d$-level spin qudit in the form
@@
 \end{align}
-Given empirical estimates $E_j^V$ of the expectation values $\bk{\Pi_j^V}_\rho$, an empirical estimate $\tilde\rho$ of $\rho$ is then
+Given empirical estimates $E_j$ of the expectation values $\bk{T_j}_\rho$, an empirical estimate $\tilde\rho$ of $\rho$ is then
 \begin{align}
@@
 \begin{align}
-  \tilde\rho^V_k \equiv \sum_j g^V_{kj} E_j^V
-  \approx \sum_j g^V_{kj} \bk{\Pi_j^V}_\rho
+  \tilde\rho^V_k \equiv \sum_j g^V_{kj} E_j
+  \approx \sum_j g^V_{kj} \bk{T_j}_\rho
   = \bk{Q^V_k}_\rho
@@
 \end{align}
-The eigenvalues and eigenvectors of the overlap matrix $G_V$ provide strict upper bounds on the distance between the empirical estimate $\tilde\rho$ and the true state $\rho$.
-If the estimates $E^V_j$ approximate $\bk{\Pi^V_j}_\rho$ to within an error $\epsilon$, i.e.
+The eigenvalues of the overlap matrix $G_V$ allow us to make concrete statements about the statistical error between the empirical estimate $\tilde\rho$ and the true state $\rho$.
+Assume, for example, that the estimates $E_j$ are equal to $\bk{T_j}_\rho$ up to Gaussian noise with variance at most $\epsilon^2$, i.e.
 \begin{align}
-  \abs{E^V_j-\bk{\Pi^V_j}_\rho} \le \epsilon,
+  E_j = \bk{T_j}_\rho + \epsilon_j,
+  &&
+  \bbk{\epsilon_j \epsilon_{j'}} \le \epsilon^2 \delta_{jj'},
 \end{align}
-then the error in the estimate $\tilde\rho^V_k$ of $\rho^V_k$ is
+where $\set{\epsilon_j}$ are independent random variables, and we use the double brackets $\bbk{\cdot}$ to denote statistical averaging.
+The variance in the error with which $\tilde\rho^V_k$ approximates $\rho^V_k$ is then
 \begin{align}
-  \abs{\tilde\rho^V_k-\rho^V_k}
-  &= \abs{\sum_j g_{kj}^V \p{E^V_j-\bk{\Pi^V_j}_\rho}} \\
-  &\le \sum_j \abs{g_{kj}^V \p{E^V_j-\bk{\Pi^V_j}_\rho}} \\
-  &\le \epsilon \norm{\v g^V_k}_1,
+  \Bbk{\abs{\tilde\rho^V_k-\rho^V_k}^2}
+  &= \Bbk{\p{\tilde\rho^V_k-\rho^V_k}^\dag \p{\tilde\rho^V_k-\rho^V_k}} \\
+  &= \sum_{j,j'} \p{g^V_{kj}}^* g^V_{kj'}\,
+  \bbk{\epsilon_j \epsilon_{j'}} \\
+  &\le \sum_j \abs{g^V_{kj}}^2 \epsilon^2
+  = \f{\epsilon^2}{G^V_k}.
 \end{align}
-where $\norm{\v g}_1=\sum_j\abs{g_j}$ is the 1-norm of the vector $\v g$.
-Using the fact that the operators $Q^V_k$ are orthonormal, we can bound the 2-norm (Euclidean) distance $\norm{\tilde\rho-\rho}_2$ between $\tilde\rho$ and $\rho$ as
+Using the fact that the operators $Q^V_k$ are orthonormal, we can therefore bound the mean squared 2-norm (Euclidean) distance between $\tilde\rho$ and $\rho$ as
 \begin{align}
-  \norm{\tilde\rho-\rho}_2^2
-  = \sum_k \abs{\tilde\rho^V_k - \rho^V_k}^2
+  \Bbk{\norm{\tilde\rho-\rho}_2^2}
+  = \sum_k \Bbk{\abs{\tilde\rho^V_k - \rho^V_k}^2}
   \le \epsilon^2 \N_V^2,
@@
 \begin{align}
-  \N_V^2 \equiv \min_U \sum_k \norm{U\v g^V_k}_1^2
-  = \min_U \sum_k \f{\norm{U\v x^V_k}_1^2}{G^V_k}
+  \N_V^2 \equiv \sum_k \f1{G^V_k}
   \label{eq:norm}
 \end{align}
-is a norm uniquely determined by the set of measurement axes $V$, with minimization performed over unitary transformations $U$ of all $\v g^V_k$ and $\v x^V_k$ to account for the fact that their 1-norms are generally basis-dependent.
-Seen otherwise, this minimization accounts for the fact that we can repeat the analysis between Eqs.~\eqref{eq:overlap_matrix} and \eqref{eq:bound} with any $G_V\to U G_V U^\dag$, where $U$ is an arbitrary unitary that mixes the projectors in $\v\Pi_V$.
-[\red{Actually, I don't think this minimization is quite correct as stated, but it won't matter after some pending changes to this section.
-Rather than an absolute error $\epsilon$, we can assume that $E_j$ approximates $\bk{\Pi_j}_\rho$ up to Gaussian noise with variance $\epsilon^2$.
-In that case, I think the mean (squared) distance between $\tilde\rho$ and $\rho$ is $\epsilon^2\sum_k1/G^V_k$, which in retrospect is a much nicer result.
-I am still working to write this up, which will involve some changes to this section.}]
+is a norm uniquely determined by the set of measurement axes $V$.
+Note that the sum in Eq.~\eqref{eq:norm} is performed over the $d^2$ largest eigenvalues of $G_V$, and therefore may be divergent if $G_V$ has fewer than $d^2$ nonzero eigenvalues.
+Altogether, diagonalizing each of $d$ blocks of $G_V$ takes $O(\abs{V}^3)$ time, so the norm $\N_V$ can be computed in $O(d\abs{V}^3)$ serial or $O(\abs{V}^3)$ parallel runtime.
 
-To better understand the bound in Eq.~\eqref{eq:bound}, we observe that by a change of operator basis, or equivalently a particular choice of unitary $U_T$, measuring the projectors $\set{\Pi_{\v v\mu}}$ is equivalent to measuring the transition operators $\set{T_{\v v\ell,0}}$.
-Orthogonality of transition operators with different degrees $\ell$ then implies that the overlap matrix
-\begin{align}
-  G_V^{\t{trn}} \equiv U_T G_V U_T^\dag
-  = \sum_{\ell,\v v,\v v'} \obk{T_{\v v\ell,0}|T_{\v v'\ell,0}}
-  \op{\v v\ell}{\v v'\ell}
-\end{align}
-has a block-diagonal structure, with $d$ blocks of size $\abs{V}\times\abs{V}$.
-Eigenvectors of $G_V$ can therefore have at most $\abs{V}$ nonzero components in the transition operator basis, which implies that
-\begin{align}
-  \N_V \le \sum_k \f{\norm{U_T\v x^V_k}_1^2}{G^V_k}
-  \le \sum_k \f{\abs{V}\,\norm{U_T\v x^V_k}_2^2}{G^V_k}
-  = \sum_k \f{\abs{V}}{G^V_k}.
-\end{align}
-Note that these sums are performed over the $d^2$ largest eigenvalues of $G_V$ (or $G_V^{\t{trn}}$), and therefore may be divergent if $G_V$ has fewer than $d^2$ nonzero eigenvalues.
-Altogether, diagonalizing each of $d$ blocks of $G_V^{\t{trn}}$ takes a time that is $O(\abs{V}^3)$, so an upper bound for $\N_V$ can be computed in $O(d\abs{V}^3)$ serial or $O(\abs{V}^3)$ parallel runtime.
-
-The bound in Eq.~\eqref{eq:bound} motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and variationally optimize the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the cost function
-\begin{align}
-  \tilde\N_V^2 \equiv \sum_k \f{\norm{\v y^V_k}_1^2}{G^V_k}
-  \ge \N_V^2,
-\end{align}
-where $\v y^V_k$ are the (normalized) eigenvectors of the overlap matrix $G_V^{\t{trn}}$ in the transition operator basis, with corresponding eigenvalues $G^V_k$.
-If $\abs{V}$ is too large for variational optimization to be practical, one can instead generate many random sets of measurement axes, and then choose the set that minimizes $\tilde\N_V^2$.
-Upon choosing a set of measurement axes $V$, experimentally characterize the expectation values $\bk{\Pi^V_j}_\rho$.
-If these measurements can approximate the expectation values $\bk{\Pi^V_j}_\rho$ to within a known error $\epsilon$, then the empirical estimate $\tilde\rho$ defined in Eq.~\eqref{eq:reconstructed_state} is an approximation of the true state $\rho$ to within a 2-norm distance of $\epsilon\N_V$.
-If the empirical estimate $\tilde\rho$ has negative eigenvalues, its 2-norm distance from $\rho$ can be further reduced via maximum-likelihood corrections \cite{smolin2012efficient}.
-In terms of classical computing costs before and after estimating the expectation values $\bk{\Pi^V_j}_\rho$, this tomography protocol is bottlenecked by the need to diagonalize the overlap matrix $G_V^{\t{trn}}$, which has $d$ blocks of size $\abs{V}\times\abs{V}$.
+The bound in Eq.~\eqref{eq:bound} motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and variationally optimize the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the error norm $\N_V$.
+If $\abs{V}$ is too large for variational optimization to be practical, one can instead generate many random sets of measurement axes, and then choose the set that minimizes $\N_V$.
+After choosing a set of measurement axes $V$, experimentally characterize the expectation values $\bk{T_j}_\rho$.
+If the experimental measurements $E_j$ of $\bk{T_j}_\rho$ are accurate to within an error $\lesssim\epsilon$, then the state $\tilde\rho$ defined in Eq.~\eqref{eq:reconstructed_state} is an approximation of the true state $\rho$ to within a 2-norm distance of $\lesssim\epsilon\N_V$.
+Furthermore, if the empirical estimate $\tilde\rho$ has negative eigenvalues, its 2-norm distance from $\rho$ can be further reduced with maximum-likelihood corrections \cite{smolin2012efficient}.
+In terms of classical computing costs before and after estimating expectation values $\bk{T_j}_\rho$, this tomography protocol is bottlenecked by the need to diagonalize the overlap matrix $G_V$, which has $d$ blocks of size $\abs{V}\times\abs{V}$.
 Altogether, this protocol therefore has an $O(d\abs{V}^3)$ serial or $O(\abs{V}^3)$ parallel runtime.
@@
 
-As a last point, we mention that the cost function $\tilde\N_V^2$ can also evaluate the quality of measurement axes chosen using existing methods for sparse signal recovery on the sphere \cite{rauhut2011sparse, alem2012sparse}.
+As a last point, we mention that the cost function $\N_V$ can also evaluate the quality of measurement axes chosen using existing methods for sparse signal recovery on the sphere \cite{rauhut2011sparse, alem2012sparse}.
 In practice, we find that uniform random sampling outperforms these methods, in the sense that it generates low-cost sets of measurement axes with higher probability.

### 2020-10-28 sun_phases/spin_qudit_tomo.tex
minor wording changes

@@
   By mapping this quantum tomography problem onto the classical problem of signal recovery on the sphere, we prove that full reconstruction of arbitrary mixed states is possible, and requires a minimum number of measurement axes, $r_d$, that is bounded by $2d-1\le r_d\le d^2$.
-  We conjecture that $r_d=2d-1$, which we have verified for all $d\le120$ using numerical experiments.
-  We then provide an algorithm with $O\p{r^3}$ runtime for certifying strict upper bounds on the error with which measurements of spin projection along $r$ measurement axes can be used to reconstruct an unknown quantum state.
-  Our algorithm motivates a simple randomized tomography protocol, which we use to study the trade-off between the number of measurement axes and state reconstruction error [\red{note: this part is not yet complete}].
+  We conjecture that $r_d=2d-1$, which we have verified for all $d\le150$ using numerical experiments.
+  We then provide a parallelized algorithm with $O\p{r^3}$ runtime for certifying upper bounds on the error with which measurements of spin projection along $r$ quantization axes can be used to reconstruct an unknown quantum state.
+  Our algorithm motivates a simple randomized tomography protocol, which we use to study the trade-off between the number of measurement axes and state reconstruction error at a fixed total number of measurements [\red{note: this part is not yet complete}].
   [\red{todo: maybe look into a simplified tomography procedure for pure states?]}
@@
 where integration is performed over all $\v v\in\Omega_2\simeq[0,2\pi]\times[0,\pi]$ with integration measure $\d\Omega_2\p{\v v} \equiv \d\alpha_{\v v} \d\beta_{\v v} \sin\beta_{\v v}$, and for ease of notation we implicitly associate doublets $\v v=\p{\alpha_{\v v},\beta_{\v v}}$ with the triplet $\p{\alpha_{\v v},\beta_{\v v},0}$, so that for example $R\p{\v v} = R\p{\alpha_{\v v},\beta_{\v v},0}$ and $T_{\v v\ell m} = T_{\p{\alpha_{\v v},\beta_{\v v},0},\ell m}$.
-The same orthogonality relations in Eqs.~\eqref{eq:ortho_full} and \eqref{eq:ortho_sphere} are inherited by the rotation matrix elements $\D_{mn}^\ell$ through the fact $\R\p{\v\omega}$ is the adjoint action of $R\p{\v\omega}$ (see Appendix \ref{sec:ortho}), and the functions $\set{\D_{m,0}^\ell}$ restricted to $\Omega_2$ are similarly another orthogonal basis for the space $L^2\p{\Omega_2}$ of square-integrable functions on the sphere.
+The same orthogonality relations in Eqs.~\eqref{eq:ortho_full} and \eqref{eq:ortho_sphere} are inherited by the adjoint rotation matrix elements $\D_{mn}^\ell$ through the fact $\R\p{\v\omega}$ is the adjoint action of $R\p{\v\omega}$ (see Appendix \ref{sec:ortho}), and the functions $\set{\D_{m,0}^\ell}$ restricted to $\Omega_2$ are similarly another orthogonal basis for the space $L^2\p{\Omega_2}$ of square-integrable functions on the sphere.
 [\red{todo: make figures visualizing $Y_{\ell m}$ and $\D^\ell_{m,0}$ for comparison}]
@@
 Unfortunately, we have no general proof of this conjecture, which may require a deeper analysis of fixed-degree signal recovery on the sphere.
-Nonetheless, we have used the tomography protocol in the next section to numerically verify that $2d-1$ measurement axes suffice for all $d\le120$.
+Nonetheless, we have used the tomography protocol in the next section to numerically verify that $2d-1$ measurement axes suffice for all $d\le150$.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Optimized tomography protocols and reconstruction error bounds}
+\section{Reconstruction error bounds}
 \label{sec:protocol}
@@
 \end{align}
-where we neglect cross terms with different degrees $\ell$ because the corresponding inner products $\obk{T_{\v v\ell,0}|T_{\v v'\ell',0}}=0$ for all $\ell\ne\ell'$.
+where we neglect cross terms with different degrees $\ell$ because the corresponding inner products vanish, i.e.~$\obk{T_{\v v\ell,0}|T_{\v v'\ell',0}}=0$ for all $\ell\ne\ell'$.
 Eigenvectors of $G_V$ define linearly independent linear combinations of the operators in $\v T_V$.
 If the operators in $\v T_V$ span the entire space of operators on a $d$-level spin qudit, which is a necessary and sufficient condition for $V$ to allow for full state tomography, then $G_V$ must have exactly $d^2$ nonzero eigenvalues.
-Indexing these eigenvalues $G^V_k$ and the corresponding (normalized) eigenvectors $\v x^V_k\equiv(x^V_{k,1},x^V_{k,2},\cdots)$ by an integer $k\in\set{1,2,\cdots,\abs{V}d}$, we can construct the orthonormal operators
+Indexing these eigenvalues $G^V_k$ and the corresponding (normalized) eigenvectors $\v x^V_k\equiv(x^V_{k,1},x^V_{k,2},\cdots)$ by an integer $k\in\set{1,2,\cdots,d^2}$, we can construct the orthonormal operators
 \begin{align}
@@
 \end{align}
-where for shorthand we use a combined index $j\equiv\p{\v v_j,\ell_j}$ to specify both a measurement axis $\v v_j$ and a degree $\ell_j$, which identify the transition operator $T_j\equiv T_{\v v_j\ell_j,0}$.
+where for shorthand we use a combined index $j=\p{\v v_j,\ell_j}$ to specify both a measurement axis $\v v_j$ and a degree $\ell_j$, which identify the transition operator $T_j\equiv T_{\v v_j\ell_j,0}$.
 These operators allow us to expand any state $\rho$ of a $d$-level spin qudit in the form
@@
 where $\set{\epsilon_j}$ are independent random variables, and we use the double brackets $\bbk{\cdot}$ to denote statistical averaging.
-The variance in the error with which $\tilde\rho^V_k$ approximates $\rho^V_k$ is then
+The mean squared error with which $\tilde\rho^V_k$ approximates $\rho^V_k$ is then
 \begin{align}
   \Bbk{\abs{\tilde\rho^V_k-\rho^V_k}^2}
-  &= \Bbk{\p{\tilde\rho^V_k-\rho^V_k}^\dag \p{\tilde\rho^V_k-\rho^V_k}} \\
+  &= \Bbk{\p{\tilde\rho^V_k-\rho^V_k}^*\p{\tilde\rho^V_k-\rho^V_k}} \\
   &= \sum_{j,j'} \p{g^V_{kj}}^* g^V_{kj'}\,
@@
 \end{align}
-is a norm uniquely determined by the set of measurement axes $V$.
+is a norm determined by the set of measurement axes $V$.
 Note that the sum in Eq.~\eqref{eq:norm} is performed over the $d^2$ largest eigenvalues of $G_V$, and therefore may be divergent if $G_V$ has fewer than $d^2$ nonzero eigenvalues.
@@
 
-The bound in Eq.~\eqref{eq:bound} motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and variationally optimize the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the error norm $\N_V$.
+The bound in Eq.~\eqref{eq:bound} motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and variationally optimize the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the error norm $\N_V$.
 If $\abs{V}$ is too large for variational optimization to be practical, one can instead generate many random sets of measurement axes, and then choose the set that minimizes $\N_V$.
@@
 
-As a last point, we mention that the cost function $\N_V$ can also evaluate the quality of measurement axes chosen using existing methods for sparse signal recovery on the sphere \cite{rauhut2011sparse, alem2012sparse}.
+As a last point, we mention that the cost function $\N_V$ can also evaluate the quality of measurement axes chosen with existing methods for sparse signal recovery on the sphere \cite{rauhut2011sparse, alem2012sparse}.
 In practice, we find that uniform random sampling outperforms these methods, in the sense that it generates low-cost sets of measurement axes with higher probability.
-Two observations may explain the poor performance of the sparse signal recovery methods.
+Two observations may explain the poor performance of these sparse signal recovery methods.
 First, spin qudit tomography essentially boils down to recovering sparse signals with a particular harmonic structure, namely with a fixed degree $\ell$.
@@
 \begin{align}
-  \abs{\Omega_3} \equiv \int_{\Omega_3} d\Omega_3\p{\v\omega} = 8\pi^2.
+  \abs{\Omega_3} \equiv \int_{\Omega_3} \d\Omega_3\p{\v\omega} = 8\pi^2.
 \end{align}
@@
   = \f{4\pi}{2\ell+1} \delta_{\ell\ell'} \delta_{mm'},
+  \label{eq:ortho_sphere_apndx}
 \end{align}
@@
 where $f_{\ell m}\in\CC$ is a coefficient and $\set{B_{\ell m}}$ is a complete basis for $L^2\p{\Omega_2}$.
-Without loss of generality, we will assume that for a fixed degree $\ell$ the functions $\set{B_{\ell m}}$ span the degree-$\ell$ subspace of $L^2\p{\Omega_2}$, and that these functions satisfy orthogonality relations analogous to that in Eq.~\eqref{eq:ortho_sphere}.
+Without loss of generality, we will assume that for a fixed degree $\ell$ the functions $\set{B_{\ell m}}$ span the degree-$\ell$ subspace of $L^2\p{\Omega_2}$, and that these functions satisfy orthogonality relations analogous to those in Eqs.~\eqref{eq:ortho_sphere} and \eqref{eq:ortho_sphere_apndx}.
 A solution to the signal recovery problem amounts to a set of points $V=\set{\v v}\subset\Omega_2$ and coefficients $C_{\ell m}\p{\v v}$ for which
@@
 \end{align}
-Consider now a different basis $\set{\tilde B_{\ell m}}$ for $L^2\p{\Omega_2}$, with which
+Consider now a different basis $\set{\tilde B_{\ell m}}$ for $L^2\p{\Omega_2}$, satisfying the same general assumptions that we made for $\set{B_{\ell m}}$, with which
 \begin{align}
-  f\p{\v v} = \sum_{\ell, m} \tilde f_{\ell m} \tilde B_{\ell m},
+  f\p{\v v} = \sum_{\ell, m} \tilde f_{\ell m} \tilde B_{\ell m}.
 \end{align}
-where we make the same general assumptions about $\set{\tilde B_{\ell m}}$ that we made for $\set{B_{\ell m}}$.
 Substituting the harmonic expansion of $f\p{\v v}$ in the original basis $\set{B_{\ell m}}$, we find that

### 2020-10-28 sun_phases/sun_dynamics_notes.tex
changed some notation, added notes on a simple model with two large spins

@@
 
-\usepackage{mathtools} % for \coloneqq
-
 \newcommand{\up}{\uparrow}
@@
 \newcommand{\Z}{\text{Z}}
-\newcommand{\BTBX}{\text{BTB-X}}
 
@@
 
-\title{SU($n$) spin model dynamics}%
+\title{SU($d$) spin model dynamics}%
 \author{Michael A. Perlin}%
@@
 
-Multilevel atoms on an optical lattice can be used to implement a variety of fermionic and spin models with SU($n$) symmetries.
+Multilevel atoms on an optical lattice can be used to implement a variety of fermionic and spin models with SU($d$) symmetries.
 Not only are these systems powerful quantum simulators, they are also promising platforms for quantum information processing and metrology.
 Two-level atoms with SU(2)-symmetric interactions, for example, can be used to simulate dynamical phases of BCS superconductors, or to prepare entangled states that enable surpassing classical limits on measurement precision.
-When atoms have $n$ nuclear spin sublevels, inter-atomic interactions can exhibit an SU($n$) symmetry that has yet to be explored as a scientific resource.
+When atoms have $d$ nuclear spin sublevels, inter-atomic interactions can exhibit an SU($d$) symmetry that has yet to be explored as a scientific resource.
 For this reason, we wish to study the dynamical quantum phases that can be implemented with these atoms.
@@
 
-Our starting point is a simple 1D model of $n$-level fermions with nearest-neighbor tunneling and on-site (collisional) exchange interactions:
+Our starting point is a simple 1D model of $d$-level fermions with nearest-neighbor tunneling and on-site (collisional) exchange interactions:
 \begin{align}
@@
 \end{align}
-where $c_{j\mu}$ annihilates a fermionic atom on lattice site $j$ with nuclear spin projection $\mu\in\set{I,I-1,\cdots,-I}$ onto a quantization axis; $I\equiv\frac{n-1}{2}$ is the total nuclear spin of each atom (e.g.~$I=9/2$ in the case of Sr-87 with $n=10$ nuclear spin levels); $J$ is the nearest-neighbor tunneling rate; and $U$ is the two-body on-site interaction strength.
+where $c_{j\mu}$ annihilates a fermionic atom on lattice site $j$ with nuclear spin projection $\mu\in\set{I,I-1,\cdots,-I}$ onto a quantization axis; $I\equiv\frac{d-1}{2}$ is the total nuclear spin of each atom (e.g.~$I=9/2$ in the case of Sr-87 with $d=10$ nuclear spin levels); $J$ is the nearest-neighbor tunneling rate; and $U$ is the two-body on-site interaction strength.
 In terms of quasi-momentum modes, the same Hamiltonian reads
@@
   \includegraphics{sun_dispersion.pdf}
-  \caption{Single-particle dispersion relations $E_{q\mu}=-2J\cos\p{q+\mu\phi}$ for the case of $n=6$ (nuclear spin $I=5/2$) and different spin-orbit coupling angles $\phi$.
+  \caption{Single-particle dispersion relations $E_{q\mu}=-2J\cos\p{q+\mu\phi}$ for the case of $d=6$ (nuclear spin $I=5/2$) and different spin-orbit coupling angles $\phi$.
     Legend indicates the spin projection $\mu$ onto the $z$.
-    Note that these dispersion relations are ``maximally spread'' when $\phi=2\pi/n$ (as with $\phi=\pi/3$ above), and that some of these dispersions overlap (implying degeneracies at all $q$) when $\phi=2\pi/m$ for any $m\in\set{1,2,\cdots,n-1}$.}
+    Note that these dispersion relations are ``maximally spread'' when $\phi=2\pi/n$ (as with $\phi=\pi/3$ above), and that some of these dispersions overlap (implying degeneracies at all $q$) when $\phi=2\pi/m$ for any $m\in\set{1,2,\cdots,d-1}$.}
   \label{fig:disps}
@@
 where we include the identity terms for comparison with \eqref{eq:spin_pair}.
-If $n$ is arbitrary but $I\phi\ll1$, then similarly
+If $d$ is arbitrary but $I\phi\ll1$, then similarly
 \begin{align}
@@
 
-Needless to say, the single-body terms in \eqref{eq:spin_pair} explicitly break the SU($n$) symmetry of the interactions.
+Needless to say, the single-body terms in \eqref{eq:spin_pair} explicitly break the SU($d$) symmetry of the interactions.
 Even so, the fact that all single-body terms are diagonal with respect to nuclear spin projection $\mu$ implies that the overall Hamiltonian conserves all collective spin operators of the form $S_{\mu\mu}\equiv\sum_q s_{\mu\mu q}$, as well as all products thereof, which is equivalent to conserving all powers of $S_\z$ in the case of SU(2)\footnote{This list of conserved quantities is not exhaustive, and in particular a specific choice of initial state may result in additional conserved quantities.}.
 The fact that the Hamiltonian is diagonal in the ``computational'' basis of all spins (i.e.~the basis of their spin projection onto the $z$ axis) and suggests that the spin Hamiltonian in \eqref{eq:spin_pair} might be integrable, although the precise form of an appropriate Bethe ansatz or Lax formulation is unclear.
-When $\phi\ll1$ or $\phi=\pi$, however, the single-body terms of the spin Hamiltonian are built from a single spin operator (namely, $s_\z$ or $s_\a$), so extending the Bethe ansatz or Lax formulation of the SU(2) case to SU($n$) is relatively straightforward.
+When $\phi\ll1$ or $\phi=\pi$, however, the single-body terms of the spin Hamiltonian are built from a single spin operator (namely, $s_\z$ or $s_\a$), so extending the Bethe ansatz or Lax formulation of the SU(2) case to SU($d$) is relatively straightforward.
 
-It is worth mentioning at this point that we can, in principle, perform full tomography on the average single-body reduced density operator $\bar\rho$, which is equivalent to measuring all components of the mean collective spin matrix $\bk{\v S}\equiv\sum_{\mu,\nu}\bk{S_{\mu\nu}}\op{\mu}{\nu}$ (see Appendix \ref{sec:tomography} for additional details)\footnote{Fluctuations in $\v S$, and thereby $\bk{\v S\c\v S}$, are also experimentally accessible in principle, but this quantity is not accessible via mean-field theory.}.
+It is worth mentioning at this point that we can, in principle, perform full tomography on the average single-body reduced density operator $\bar\rho$, which is equivalent to measuring all components of the mean collective spin matrix $\bk{\v S}\equiv\sum_{\mu,\nu}\bk{S_{\mu\nu}}\op{\mu}{\nu}$ (we can provide a different set of notes with additional details upon request)\footnote{Fluctuations in $\v S$, and thereby $\bk{\v S\c\v S}$, are also experimentally accessible in principle, but this quantity is not accessible via mean-field theory.}.
 The squared mean ``magnetization'' $\bk{\v S}\c\bk{\v S}=\sum_{\mu,\nu}\abs{\bk{S_{\mu\nu}}}^2$, which is essentially equal to the purity of $\bar\rho$, is therefore an experimentally accessible  ``unbiased'' dynamical order parameter that we can consider.
@@
 
-The simplest initial state we can prepare is that with all atoms occupying a single nuclear spin state, e.g.~the state $\ket{\Z}\equiv\ket{I}^{\otimes N}$ with all $N$ atoms pointing up along the $z$ axis.
+The simplest initial state we can prepare is that with all atoms occupying a single nuclear spin state, e.g.~the state $\ket{\Z}\equiv\ket{I}^{\otimes N}$ with all $d$ atoms pointing up along the $z$ axis.
 Using external control fields (discussed in the Section \ref{sec:controls}), we can rotate the state $\ket\Z$ to orient it along any axis in 3D space.
 External control fields also allow us to prepare product states in which each spin is in a coherent superposition of states polarized along different directions, e.g.~$\p{\ket{+I}+\ket{-I}}^{\otimes N}$.
-Generally speaking, the possibilities to prepare $N$-fold product states of the form $\ket\psi^{\otimes N}$ are enormous.
+Generally speaking, the possibilities to prepare $d$-fold product states of the form $\ket\psi^{\otimes N}$ are enormous.
 
 Perhaps more interestingly, we can in principle prepare inhomogeneous product states in which the state of each spin is correlated with its corresponding single-body term of the spin Hamiltonian in \eqref{eq:spin_pair}.
-We are currently working on characterizing the sort of inhomogeneous states that we can prepare (see Appendix \ref{sec:init_states}).
+We are still working on characterizing the sort of inhomogeneous states that we can prepare.
 
@@
 In order to explore the entirety of parameter space, however, we turn to mean-field theory.
-To this end, we first decompose individual spin operators into Schwinger bosons, $s_{\mu\nu q}\coloneqq b_{\mu q}^\dag b_{\nu q}$.
+To this end, we first decompose individual spin operators into Schwinger bosons, $s_{\mu\nu q}\to b_{\mu q}^\dag b_{\nu q}$.
 Up to an overall constant, our spin model can then be written in the form
@@
 Our mean-field theory then simply treats the boson operators $b_{\mu q}$ as complex numbers, with the initial value $b_{\mu q}$ equal to the amplitude of spin $q$ in state $\mu$\footnote{This mean-field theory is mathematically equivalent to a mean-field treatment of spin operators $s_{\mu\nu k}$ with initial conditions set by the corresponding initial-state expectation values.
-The advantage of the Schwinger boson decomposition is that it reduces the number of variables (operators) from $n^2 N$ to $nN$.}.
+The advantage of the Schwinger boson decomposition is that it reduces the number of variables (operators) from $d^2 N$ to $dN$.}.
 
@@
 We will generally assume unit filling, setting $L=N$, and here consider spin-orbit coupling angles $\phi\in\set{\pi/2,\pi}$\footnote{In retrospect, we note that these choices of spin-orbit coupling angles are not ideal for examining ``generic'' behavior, because they yield degenerate dispersions, as discussed in the caption of Figure \ref{fig:disps}.}.
-We will also consider two types of initial states: a fully polarized state $\ket\X$ in which all spins point along the $x$ axis, and a ``back-to-back'' state $\ket\BTBX$ in which half of the spins are polarized along $+x$, and half along $-x$.
-Specifically, indexing spins by their associated quasi-momentum mode $q\in(-\pi,\pi)$, the initial state $\ket\BTBX$ has spins with $q>0$ ($q<0$) initially polarized along $+x$ ($-x$).
-In this way, the initial state $\ket\BTBX$ is correlated with $\sim\sin q$ terms of the spin Hamiltonian in \eqref{eq:spin_pair}.
+We will also consider two types of initial states: a fully polarized state $\ket\X$ in which all spins point along the $x$ axis, and a ``back-to-back'' state $\ket{\X\X}$ in which half of the spins are polarized along $+x$, and half along $-x$.
+Specifically, indexing spins by their associated quasi-momentum mode $q\in(-\pi,\pi)$, the initial state $\ket{\X\X}$ has spins with $q>0$ ($q<0$) initially polarized along $+x$ ($-x$).
+In this way, the initial state $\ket{\X\X}$ is correlated with $\sim\sin q$ terms of the spin Hamiltonian in \eqref{eq:spin_pair}.
 
-For each initial state, $\ket\X$ or $\ket\BTBX$, and each spin-orbit coupling angle $\phi\in\set{\pi/2,\pi}$, we will examine how the dynamical behavior of a few quantities depends on the strength of single-particle terms relative to interactions, $J/U$.
+For each initial state, $\ket\X$ or $\ket{\X\X}$, and each spin-orbit coupling angle $\phi\in\set{\pi/2,\pi}$, we will examine how the dynamical behavior of a few quantities depends on the strength of single-particle terms relative to interactions, $J/U$.
 In particular, we will examine the mean squared magnetization
@@
 
-Figures \ref{fig:ss_spect_n2_X_p0.5}--\ref{fig:ss_spect_n2_XX_p1.0} show the time-series of $s^2(t)$, its power spectral density $\widetilde{s^2}(\omega)$, and the time-series spectrum of $\bar\rho(t)$ for the case of $n=2$, with values of $J/U$ between $10^{-2}$ and $10^1$.
+Figures \ref{fig:ss_spect_n2_X-L_a0.5}--\ref{fig:ss_spect_n2_XX-L_a1.0} show the time-series of $s^2(t)$, its power spectral density $\widetilde{s^2}(\omega)$, and the time-series spectrum of $\bar\rho(t)$ for the case of $d=2$, with values of $J/U$ between $10^{-2}$ and $10^1$.
 In short, these figures all tell the same story: there is a ``collective'' phase when $J/U\lesssim1$, and of a ``single-particle'' phase when $J/U\gtrsim1$.
@@
 
-Figure \ref{fig:ss_spect_n6_X_p0.5}--\ref{fig:ss_spect_n6_XX_p1.0} show the same plots for the case of $n=6$.
+Figure \ref{fig:ss_spect_n6_X-L_a0.5}--\ref{fig:ss_spect_n6_XX-L_a1.0} show the same plots for the case of $d=6$.
 Unsurprisingly, here again there appear to be (at least) two phases, at $J/U\lesssim1$ vs.~$J/U\gtrsim1$.
-An interesting observation, however, is that (for both initial states we consider) the spectrum of $\bar\rho$ appears to have only four non-zero eigenvalues when $\phi=\pi/2$, and two non-zero eigenvalues when $\phi=\pi$.
-The case of $\phi=\pi$ happens to be special, as it turns out that (for the initial states we consider) it can be mapped exactly onto a spin model with $n=2$, regardless of the actual value of $n$.
-The general case of $\phi\ne\pi$, however, seems to yield four eigenvalues independent of the actual value of $\phi$.
-One might suspect then that $\bar\rho$ is effectively a density operator on some fixed (time-independent) four-dimensional subspace of $\H_n$ (i.e.~the Hilbert space of an $n$-level quantum system), but that does not appear to be the case: decomposing $\bar\rho\equiv\sum_\lambda\lambda\op\lambda$, we find that the projector $\P_{\bar\rho}\equiv\sum_{\lambda>0}\op\lambda$ is time dependent\footnote{To clarify: $\P_{\bar\rho}$ is time dependent even if you only consider times when $\tr\P_{\bar\rho}=4$.
+An interesting observation, however, is that the spectrum of $\bar\rho$ appears to have only four non-zero eigenvalues when $\phi=\pi/2$, and two non-zero eigenvalues when $\phi=\pi$.
+The case of $\phi=\pi$ happens to be special, as it turns out that (for the initial states we consider) it can be mapped exactly onto a spin model with $d=2$, regardless of the actual value of $d$.
+The general case of $\phi\ne\pi$, however, seems to yield four eigenvalues independent of the value of $\phi$.
+One might suspect then that $\bar\rho$ is effectively a density operator on some fixed (time-independent) four-dimensional subspace of $\H_n$ (i.e.~the Hilbert space of an $d$-level quantum system), but that does not appear to be the case: decomposing $\bar\rho\equiv\sum_\lambda\lambda\op\lambda$, we find that the projector $\P_{\bar\rho}\equiv\sum_{\lambda>0}\op\lambda$ is time dependent\footnote{To clarify: $\P_{\bar\rho}$ is time dependent even if you only consider times when $\tr\P_{\bar\rho}=4$.
   It should be possible to understand this fact in the non-interacting limit $U/J\to0$.}!
@@
 \begin{figure}
-  \centering \includegraphics{oscillations/time_series/ss_spect_n2_N20_X_p0.50.pdf}
-  \caption{Time-series and power spectral density of $s^2$ (left and middle columns), as well as the spectrum of $\bar\rho$ (right column), for $N=20$ spins with $n=2$ initially in $\ket\X$, with spin-orbit coupling angle $\phi=\pi/2$.}
-  \label{fig:ss_spect_n2_X_p0.5}
+  \centering
+  \includegraphics{oscillations/ss_spect_n2_N20_X-L_a0.50.pdf}
+  \caption{Time-series and power spectral density of $s^2$ (left and middle columns), as well as the spectrum of $\bar\rho$ (right column), for $N=20$ spins with $d=2$ initially in $\ket\X$, with spin-orbit coupling angle $\phi=\pi/2$.}
+  \label{fig:ss_spect_n2_X-L_a0.5}
 \end{figure}
@@
   \centering
-  \includegraphics{oscillations/time_series/ss_spect_n2_N20_X_p1.00.pdf}
-  \caption{Time-series and power spectral density of $s^2$ (left and middle columns), as well as the spectrum of $\bar\rho$ (right column), for $N=20$ spins with $n=2$ initially in $\ket\X$, with spin-orbit coupling angle $\phi=\pi$.}
-  \label{fig:ss_spect_n2_X_p1.0}
+  \includegraphics{oscillations/ss_spect_n2_N20_X-L_a1.00.pdf}
+  \caption{Time-series and power spectral density of $s^2$ (left and middle columns), as well as the spectrum of $\bar\rho$ (right column), for $N=20$ spins with $d=2$ initially in $\ket\X$, with spin-orbit coupling angle $\phi=\pi$.}
+  \label{fig:ss_spect_n2_X-L_a1.0}
 \end{figure}
@@
 \begin{figure}
-  \centering \includegraphics{oscillations/time_series/ss_spect_n2_N20_XX_p0.50.pdf}
-  \caption{Time-series and power spectral density of $s^2$ (left and middle columns), as well as the spectrum of $\bar\rho$ (right column), for $N=20$ spins with $n=2$ initially in $\ket\BTBX$, with spin-orbit coupling angle $\phi=\pi/2$.}
-  \label{fig:ss_spect_n2_XX_p0.5}
+  \centering
+  \includegraphics{oscillations/ss_spect_n2_N20_XX-L_a0.50.pdf}
+  \caption{Time-series and power spectral density of $s^2$ (left and middle columns), as well as the spectrum of $\bar\rho$ (right column), for $N=20$ spins with $d=2$ initially in $\ket{\X\X}$, with spin-orbit coupling angle $\phi=\pi/2$.}
+  \label{fig:ss_spect_n2_XX-L_a0.5}
 \end{figure}
@@
   \centering
-  \includegraphics{oscillations/time_series/ss_spect_n2_N20_XX_p1.00.pdf}
-  \caption{Time-series and power spectral density of $s^2$ (left and middle columns), as well as the spectrum of $\bar\rho$ (right column), for $N=20$ spins with $n=2$ initially in $\ket\BTBX$, with spin-orbit coupling angle $\phi=\pi$.}
-  \label{fig:ss_spect_n2_XX_p1.0}
+  \includegraphics{oscillations/ss_spect_n2_N20_XX-L_a1.00.pdf}
+  \caption{Time-series and power spectral density of $s^2$ (left and middle columns), as well as the spectrum of $\bar\rho$ (right column), for $N=20$ spins with $d=2$ initially in $\ket{\X\X}$, with spin-orbit coupling angle $\phi=\pi$.}
+  \label{fig:ss_spect_n2_XX-L_a1.0}
 \end{figure}
@@
 \begin{figure}
-  \centering \includegraphics{oscillations/time_series/ss_spect_n6_N20_X_p0.50.pdf}
-  \caption{Time-series and power spectral density of $s^2$ (left and middle columns), as well as the spectrum of $\bar\rho$ (right column), for $N=20$ spins with $n=6$ initially in $\ket\X$, with spin-orbit coupling angle $\phi=\pi/2$.}
-  \label{fig:ss_spect_n6_X_p0.5}
+  \centering
+  \includegraphics{oscillations/ss_spect_n6_N20_X-L_a0.50.pdf}
+  \caption{Time-series and power spectral density of $s^2$ (left and middle columns), as well as the spectrum of $\bar\rho$ (right column), for $N=20$ spins with $d=6$ initially in $\ket\X$, with spin-orbit coupling angle $\phi=\pi/2$.}
+  \label{fig:ss_spect_n6_X-L_a0.5}
 \end{figure}
@@
   \centering
-  \includegraphics{oscillations/time_series/ss_spect_n6_N20_X_p1.00.pdf}
-  \caption{Time-series and power spectral density of $s^2$ (left and middle columns), as well as the spectrum of $\bar\rho$ (right column), for $N=20$ spins with $n=6$ initially in $\ket\X$, with spin-orbit coupling angle $\phi=\pi$.}
-  \label{fig:ss_spect_n6_X_p1.0}
+  \includegraphics{oscillations/ss_spect_n6_N20_X-L_a1.00.pdf}
+  \caption{Time-series and power spectral density of $s^2$ (left and middle columns), as well as the spectrum of $\bar\rho$ (right column), for $N=20$ spins with $d=6$ initially in $\ket\X$, with spin-orbit coupling angle $\phi=\pi$.}
+  \label{fig:ss_spect_n6_X-L_a1.0}
 \end{figure}
@@
 \begin{figure}
-  \centering \includegraphics{oscillations/time_series/ss_spect_n6_N20_XX_p0.50.pdf}
-  \caption{Time-series and power spectral density of $s^2$ (left and middle columns), as well as the spectrum of $\bar\rho$ (right column), for $N=20$ spins with $n=6$ initially in $\ket\BTBX$, with spin-orbit coupling angle $\phi=\pi/2$.}
-  \label{fig:ss_spect_n6_XX_p0.5}
+  \centering
+  \includegraphics{oscillations/ss_spect_n6_N20_XX-L_a0.50.pdf}
+  \caption{Time-series and power spectral density of $s^2$ (left and middle columns), as well as the spectrum of $\bar\rho$ (right column), for $N=20$ spins with $d=6$ initially in $\ket{\X\X}$, with spin-orbit coupling angle $\phi=\pi/2$.}
+  \label{fig:ss_spect_n6_XX-L_a0.5}
 \end{figure}
@@
   \centering
-  \includegraphics{oscillations/time_series/ss_spect_n6_N20_XX_p1.00.pdf}
-  \caption{Time-series and power spectral density of $s^2$ (left and middle columns), as well as the spectrum of $\bar\rho$ (right column), for $N=20$ spins with $n=6$ initially in $\ket\BTBX$, with spin-orbit coupling angle $\phi=\pi$.}
-  \label{fig:ss_spect_n6_XX_p1.0}
+  \includegraphics{oscillations/ss_spect_n6_N20_XX-L_a1.00.pdf}
+  \caption{Time-series and power spectral density of $s^2$ (left and middle columns), as well as the spectrum of $\bar\rho$ (right column), for $N=20$ spins with $d=6$ initially in $\ket{\X\X}$, with spin-orbit coupling angle $\phi=\pi$.}
+  \label{fig:ss_spect_n6_XX-L_a1.0}
 \end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{A simple model}
+
+Here we consider a relatively simple Hamiltonian on $2N$ spins,
+\begin{align}
+  H = \f{U}{2} \v S\c\v S + \sum_{j=1}^{2N} h_j s_{\z,j},
+  &&
+  h_j \equiv
+  \begin{cases}
+    +1 & j \le N \\
+    -1 & j > N
+  \end{cases}.
+  \label{eq:two_spins_start}
+\end{align}
+Defining collective the spin-$\alpha$ operators for each sub-ensemble indexed by $k\in\set{0,1}$,
+\begin{align}
+  S_{\alpha k} = \sum_{j=1+kN}^{N+kN} s_{\alpha j},
+\end{align}
+as long as each sub-ensemble remains within the manifold of spin-polarized states we can write the Hamiltonian in \eqref{eq:two_spins_start} in the form
+\begin{align}
+  H = U \v S_0 \c\v S_1 + S_{\z,0} - S_{\z,1}
+  = \sum_{m,n=-IN}^{IN} \p{U\op{mn}{nm}
+    + \p{m-n}\op{mn}},
+\end{align}
+where $m,n$ are projections of a net sub-ensemble spin onto the $z$ axis, and we have neglected the scalars $\v S_k\cdot \v S_k = IN\p{IN+1}$.
+Indexing states $\ket{mn}$ by the mean spin projection $\bar m=\p{m+n}/2$ and spin imbalance $\delta=m-n$,
+\begin{align}
+  \ket{\bar m\delta} \equiv \ket{m+\delta/2,m-\delta/2},
+\end{align}
+we can expand
+\begin{align}
+  \v S_0\c \v S_1 = \sum_{m,\delta}
+  \op{\bar m\delta}{\bar m,-\delta},
+  &&
+  S_{\z,0} - S_{\z,1} = \sum_{m,\delta} \delta \op{\bar m\delta},
+\end{align}
+which allows us to write the Hamiltonian in the form
+\begin{align}
+  H = \sum_m \op{\bar m} \otimes
+  \sum_\delta \p{U \op{\delta}{-\delta} + \delta \op{\delta}}.
+\end{align}
+Diagonalizing this Hamiltonian thus reduces to a qubit problem, for which we define the Pauli operator
+\begin{align}
+  X_\delta\p{U} \equiv \sum_{\sigma\in\set{\pm1}}
+  \sigma \op{\delta,U,\sigma},
+\end{align}
+in terms of the rotated states
+\begin{align}
+  \ket{\delta,U,\sigma} \equiv \cos\p{\f{\eta}{2}} \ket{\p{\sigma\delta}}
+  + \sigma \sin\p{\f{\eta}{2}} \ket{\p{-\sigma\delta}},
+  &&
+  \eta \equiv \arctan\p{\f{U}{\delta}},
+\end{align}
+defined for $\delta\ne0$.
+Altogether, the Hamiltonian is then
+\begin{align}
+  H = \sum_m \op{\bar m} \otimes
+  \sp{U \op{0} + \sum_{\delta>0} U_\delta X_\delta\p{U}}
+  &&
+  U_\delta \equiv \sqrt{U^2+\delta^2}.
+\end{align}
+The eigenvalues of $H$ are therefore $U$ and $\pm U_\delta$ for integer $\delta>0$.
+We now consider an initial state in which both large spins point at a polar angle $\theta$ and azimuthal angles $\pm\phi/2$:
+\begin{align}
+  \ket{\theta,\phi} \equiv \sum_{m,n}
+  \sqrt{\chi_m\p{\theta} \chi_n\p{\theta}} \,
+  e^{-\i\p{m-n}\phi} \ket{mn}
+  = \sum_{m,\delta}
+  \chi_{m\delta}\p{\theta} e^{-\i\delta\phi}
+  \ket{\bar m\delta}.
+\end{align}
+where
+\begin{align}
+  \chi_m\p\theta \equiv { 2IN \choose IN + m }
+  \cos\p{\f{\theta}{2}}^{2IN+2m} \sin\p{\f{\theta}{2}}^{2IN-2m},
+  &&
+  \chi_{m\delta}\p\theta
+  \equiv \sqrt{\chi_{m+\delta/2}\p\theta \chi_{m-\delta/2}\p\theta}.
+\end{align}
+In the eigenbasis of $H$, we have
+\begin{align}
+  \ket{\theta,\phi} = \sum_m \ket{\bar m} \otimes
+  \sp{\chi_0\p\theta \ket{0}
+    + \sum_{\delta>0} \chi_{m\delta}\p\theta
+    \sum_{\sigma\in\set{\pm1}} e^{-\i\sigma\delta\phi}
+    \sp{\cos\p{\f{\eta}{2}} \ket{\delta,U,\sigma}
+    - \sigma\sin\p{\f{\eta}{2}} \ket{\delta,U,-\sigma}}},
+\end{align}
+where again $\eta\equiv\arctan\p{U/\delta}$ for shorthand.
+Note that the only dependence of this state on the nuclear spin $I$ appears in the range of $\bar m$, so any distinction between observed behavior at different $I$ can essentially be attributed to the measurement of different observables.
+
+We now turn our attention to the measurement of collective observables.
+In particular, we consider the collective $N$-spin operators
+\begin{align}
+  S_{\mu\nu} \equiv \sum_{j=1}^N s_{\mu\nu j},
+\end{align}
+where $\mu,\nu\in\set{-I,-I+1,\cdots,I}$ and $s_{\mu\nu j} = \op{\mu}{\nu}_j$ is a spin-transition operator for spin $j$.
+Rather than initial/final spin projections $\mu\to\nu$, it will be convenient to consider operators that induce spin transitions $\mu\to\mu+\epsilon$, and then expand these operators in the Dicke basis for spin-polarized states:
+\begin{align}
+  \tilde S_{\mu\epsilon} \equiv S_{\mu+\epsilon,\mu}
+  = \sum_m \xi_{\mu\epsilon}^m \op{m+\epsilon}{m},
+  &&
+  \xi_{\mu\epsilon}^m \equiv \bk{m+\epsilon|S_{\mu+\epsilon,\mu}|m}.
+\end{align}
+In turn, we symmetrize these operators over their action on two sub-ensembles of $N$ spins each:
+\begin{align}
+  S_{\mu\epsilon\sigma} \equiv \tilde S_{\mu\epsilon} \otimes \1
+  + \sigma \1 \otimes \tilde S_{\mu\epsilon}
+  = \sum_{m,n} \xi_{\mu\epsilon}^m
+  \p{\op{m+\epsilon,n}{mn} + \sigma \op{n,m+\epsilon}{nm}},
+\end{align}
+where $\sigma\in\set{\pm1}$.
+In the symmetrized basis of a mean spin and spin imbalance, we have that
+\begin{align}
+  S_{\mu\epsilon\sigma}= \sum_{m,\delta,\sigma} \p{\xi_{\mu\epsilon}^{m+\delta/2}
+    \ket{\bar m_\epsilon,\delta+\epsilon}
+    + \sigma \xi_{\mu\epsilon}^{m-\delta/2}
+    \ket{\bar m_\epsilon,\delta-\epsilon}}
+  \bra{\bar m\delta},
+  &&
+  \bar m_\epsilon \equiv \bar m + \epsilon/2.
+\end{align}
+
+\newpage
 
@@
 \appendix
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Spin qudit tomography}
-\label{sec:tomography}
-
-Here we discuss the task of performing full tomography on the average reduced single-particle density operator of a collection of nuclear spins.
-As this task will be performed using collective measurements and homogeneous control fields, for ease of language we will consider the equivalent problem of performing tomography on a single nuclear spin.
-We have essentially two ingredients at our disposal: projective measurements of spin onto a fixed quantization axis, and a three-laser drive that addresses nuclear spins via off-resonant coupling to an excited electronic state.
-To be concrete, we can directly measure projectors $\op{\mu}$ onto states of definite spin projection $\mu$ onto the $z$ axis.
-While the three-laser drive gives us access to a variety of nuclear spin Hamiltonians, most notably it allows us to implement SU(2) rotations of the form $e^{-\i\v\theta\c\v s_{\t{3D}}}$, where $\v s_{\t{3D}}\equiv\p{s_\x,s_\y,s_\z}$ is a vector of spin operators that generate rotations of an $n$-level spin (with total spin $I=\frac{n-1}{2}$) in 3D space, and $\v\theta\equiv\p{\theta_\x,\theta_\y,\theta_\z}$ is an arbitrary rotation vector (see Appendix \ref{sec:controls}).
-These rotations essentially allow us to measure projectors $\Pi_{\v v\mu} \equiv \op{\mu_{\v v}}$ onto states of definite spin projection $\mu$ onto an arbitrary quantization axis $\v v\in\SS_2\setminus\ZZ_2$ (where we mod out by $\ZZ_2$ because antipodal points on the sphere correspond to the same set of projectors).
-
-Question: given a collection of projectors $\Pi\equiv\set{\Pi_j}$ and measurement outcomes $M\equiv\set{M_j}$, with $M_j$ an empirical estimate of $\bk{\Pi_j}_\rho=\tr\p{\rho\Pi_j}$, how do we determine whether the measurement data $M$ is sufficient to reconstruct $\rho$?
-Answer: reconstruction of $\rho$ is possible if $\Pi$ spans the entire space $\B_n$ of operators on the Hilbert space of an $n$-level spin.
-Given an orthonormal basis $\set{Q_\alpha}$ of self-adjoint operators spanning $\B_n$, we can expand
-\begin{align}
-  \rho = \sum_\alpha \rho_\alpha Q_\alpha,
-  &&
-  \rho_\alpha \equiv \bk{Q_\alpha}_\rho.
-\end{align}
-If $\Pi$ spans $\B_n$, then we can find a set of real numbers $c_{\alpha j}\in\RR$ for which
-\begin{align}
-  Q_\alpha = \sum_j c_{\alpha j} \Pi_j.
-\end{align}
-An empirical estimate $\tilde\rho$ of $\rho$ is then\footnote{This estimate can be improved by defining $\tilde\rho$ as the least-squares fit to the set of linear equations $\tr\p{\tilde\rho\,\Pi_j} = M_j$, and performing maximum-likelihood corrections to $\tilde\rho$ \cite{smolin2012efficient}.}
-\begin{align}
-  \tilde\rho = \sum_\alpha \tilde\rho_\alpha Q_\alpha,
-  &&
-  \tilde\rho_\alpha \equiv \sum_j c_{\alpha j} M_j
-  \approx \sum_j c_{\alpha j} \bk{\Pi_j}_\rho
-  = \bk{Q_\alpha}_\rho
-  = \rho_\alpha.
-\end{align}
-As it turns out, numerical experiments with $n\le20$ suggest that in order to perfurm full tomography of $\rho$, it suffices to measure all projectors $\Pi_{\v v\mu}$ onto states of definite spin $\mu$ along $2n-1$ different axes $\v v$\footnote{It would be interesting to develop an actual (analytical) understanding of this numerical discovery.}.
-The question remains: which axes should we choose?
-As a first pass, we can simply choose a set of $2n-1$ random axes, $V\equiv\set{\v v}$, by uniformly sampling points on the sphere.
-These random axes define a set of $\p{2n-1}\times n$ projectors $\Pi_V=\set{\Pi^V_j}$, where for shorthand we use a combined index $j=\p{\v v_j,\mu_j}$ that specifies both a measurement axis $\v v_j$ and an outcome $\mu_j$, and for clarity we index objects by $V$ to emphasize their dependence on the choice of measurement axes.
-The projectors $\Pi_V$ can be used to construct $n^2$ linearly independent operators by constructing a matrix $B_V$ with entries
-\begin{align}
-  \sp{B_V}_{ij} = \tr\p{\Pi^V_i \Pi^V_j}.
-\end{align}
-Indexing the (normalized) eigenvectors and the corresponding eigenvalues of $B_V$ by $\alpha$, respectively $\v x^V_\alpha$ and $m^V_\alpha$, for all $\alpha$ with eigenvalue $m^V_\alpha\ne0$ we can then construct the normalized and linearly independent operators
-\begin{align}
-  Q^V_\alpha \equiv \f1{\sqrt{m^V_\alpha}} \sum_j x^V_{\alpha j} \Pi^V_j.
-\end{align}
-To summarize: choosing a set $V$ of $2n-1$ random axes will determine a set of measurement operators $\Pi_V$ from which we can construct a complete basis $\set{Q^V_\alpha}$ of operators for $\B_n$.
-However, some choices of axes will be better than others: a good choice of axes $V$ will maximize the amount of information that the corresponding measurement data $M_V=\set{M^V_j}$ will contain about the expectation values $\bk{Q^V_\alpha}_\rho$ used to reconstruct $\rho$.
-The amount of information that $M_V$ contains about $\bk{Q^V_\alpha}_\rho$ is determined, in part, by the magnitude of the eigenvalue $m^V_\alpha$: if $m^V_\alpha$ is small, then $M_V$ will generally contain little information about $\bk{Q^V_\alpha}_\rho$, and vice versa\footnote{Think about the limit $m^V_\alpha\to0$, in which case the fact that $Q^V_\alpha$ is normalized implies that the operator $\sum_j x^V_{\alpha j} \Pi^V_j$ must have a vanishing norm.
-  A vanishing norm can only occur due to fine-tuned cancellations that make the expectation value of $\sum_j x^V_{\alpha j} \Pi^V_j$ difficult to estimate from noisy data on $\bk{\Pi^V_j}_\rho$.}.
-Rather than choosing $2n-1$ axes randomly, we can therefore choose a set of axes that minimizes some cost function $C\p{m_V}$ favoring large eigenvalues $m_V\equiv\set{m^V_\alpha}$.
-A reasonable cost function that treats all eigenvalues on equal footing, for example, is
-\begin{align}
-  C_{\t{inv}}\p{m_V} \equiv \sum_\alpha \f1{m^V_\alpha},
-\end{align}
-where the sum is implicitly performed over the $n^2$ largest eigenvalues of $B_V$ (as all other eigenvalues are guaranteed to be zero).
-Note that any information about $\rho$, obtained either from prior knowledge or preliminary measurement data, can be used to construct an even better choice of measurement axes, leading to the possibility of tailored or adaptive measurement protocols \cite{pereira2018adaptive} that are more efficient in terms of the number of measurements required to estimate $\rho$ to a fixed precision.
-We leave the details of tailored and adaptive measurement protocols to future work.
 
@@
 \end{align}
-where $s_{\v v\pm}$ are spin-raising/lowering operators along axis $\v v$ (containing all appropriate Clebsch-Gordan coefficients), and $\1$ is the identity operator for a single $n$-level spin (with total spin $I=\frac{n-1}{2}$).
-To write out $s_{\v v\pm}$ in full, we first identify the operators $s_\x,s_\y,s_\z$ that generate rotations of an $n$-level spin in 3D space, and define $s_\pm\equiv s_\x\pm\i s_\y$.
+where $s_{\v v\pm}$ are spin-raising/lowering operators along axis $\v v$ (containing all appropriate Clebsch-Gordan coefficients), and $\1$ is the identity operator for a single $d$-level spin (with total spin $I=\frac{d-1}{2}$).
+To write out $s_{\v v\pm}$ in full, we first identify the operators $s_\x,s_\y,s_\z$ that generate rotations of an $d$-level spin in 3D space, and define $s_\pm\equiv s_\x\pm\i s_\y$.
 Denoting the azimuthal and polar angles of $\v v$ respectively by $\alpha_{\v v}$ and $\beta_{\v v}$, we can use the fact that $s_\x,s_\y,s_\z$ form an $\su(2)$ algebra to expand
@@
 \end{align}
-We do not expect these controls to be sufficient for implementing arbitrary collective SU($n$) rotations, but identifying the subgroup of SU($n$) that can be generated by \eqref{eq:drive_single} could be an interesting problem for future work.
+We do not expect these controls to be sufficient for implementing arbitrary collective SU($d$) rotations, but identifying the subgroup of SU($d$) that can be generated by \eqref{eq:drive_single} could be an interesting problem for future work.
 A simple special case to consider is that of all $\eta_m=0$, in which case
@@
 One notable capability that these controls provide us is that of using pulse sequences to implement SU(2) rotations of the form $e^{-\i\v\theta\c\v s_{\t{3D}}}$, where $\v s_{\t{3D}}\equiv\p{s_\x,s_\y,s_\z}$ and $\v\theta\equiv\p{\theta_\x,\theta_\y,\theta_\z}$ is an arbitrary rotation vector.
-In addition to enabling arbitrary 3D rotations, these collective rotations enable us to measure all components of the mean collective spin matrix $\bk{\v S}$, as discussed in Appendix \ref{sec:tomography}.
+In addition to enabling arbitrary 3D rotations, these collective rotations enable us to measure all components of the mean collective spin matrix $\bk{\v S}$.
 
@@
 \end{align}
-where $\mu,\nu$ index orthogonal states of an $n$-level spin; $j,k\in\ZZ_N$ index one of $N$ spins; $h^{\mu\nu j}_{\rho\sigma k}$ and $\epsilon_{\mu\nu j}$ are scalars; and $s_{\mu\nu j}\coloneqq\op{\mu}{\nu}_j$ is a transition operator for spin $j$.
+where $\mu,\nu$ index orthogonal states of an $d$-level spin; $j,k\in\ZZ_N$ index one of $d$ spins; $h^{\mu\nu j}_{\rho\sigma k}$ and $\epsilon_{\mu\nu j}$ are scalars; and $s_{\mu\nu j}=\op{\mu}{\nu}_j$ is a transition operator for spin $j$.
 We can write this Hamiltonian using Schwinger bosons as
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Initial states}
-\label{sec:init_states}
-
-Here we explore the preparation of initial states that are correlated with the single-particle terms of the spin Hamiltonian in Eqs.~\eqref{eq:spin_cos} and \eqref{eq:spin_pair}, namely
-\begin{align}
-  H_{\t{SOC}} \equiv -2 J \sum_{q,\mu} \cos\p{q+\mu\phi} s_{\mu\mu q}.
-\end{align}
-
 \bibliography{sun_dynamics_notes.bib}

### 2020-10-28 sun_phases/spin_qudit_tomo.tex
updated verifying conjecture of r_d = 2d-1 up to d <= 200

@@
   By mapping this quantum tomography problem onto the classical problem of signal recovery on the sphere, we prove that full reconstruction of arbitrary mixed states is possible, and requires a minimum number of measurement axes, $r_d$, that is bounded by $2d-1\le r_d\le d^2$.
-  We conjecture that $r_d=2d-1$, which we have verified for all $d\le150$ using numerical experiments.
+  We conjecture that $r_d=2d-1$, which we have verified for all $d\le200$ using numerical experiments.
   We then provide a parallelized algorithm with $O\p{r^3}$ runtime for certifying upper bounds on the error with which measurements of spin projection along $r$ quantization axes can be used to reconstruct an unknown quantum state.
@@
 Unfortunately, we have no general proof of this conjecture, which may require a deeper analysis of fixed-degree signal recovery on the sphere.
-Nonetheless, we have used the tomography protocol in the next section to numerically verify that $2d-1$ measurement axes suffice for all $d\le150$.
+Nonetheless, we have used the tomography protocol in the next section to numerically verify that $2d-1$ measurement axes suffice for all $d\le200$.
 

### 2020-10-28 sun_phases/sun_dynamics_notes.tex
minor changes to the 'simple model' of two large spins

@@
 
-Here we consider a relatively simple Hamiltonian on $2N$ spins,
+\subsection{Hamiltonian}
+
+Here we consider a simple Hamiltonian on $2N$ spins,
 \begin{align}
@@
   \v S_0\c \v S_1 = \sum_{m,\delta}
-  \op{\bar m\delta}{\bar m,-\delta},
+  \op{\bar m,-\delta}{\bar m\delta},
   &&
@@
   H = \sum_m \op{\bar m} \otimes
-  \sum_\delta \p{U \op{\delta}{-\delta} + \delta \op{\delta}}.
+  \sum_\delta \p{U \op{-\delta}{\delta} + \delta \op{\delta}}.
 \end{align}
@@
 defined for $\delta\ne0$.
-Altogether, the Hamiltonian is then
+In words, the state $\ket{\delta,U,\pm}$ points in the direction of $\pm\p{\delta\uv{z}+U\uv{x}}$ on the Bloch sphere defined by the state $\set{\ket{\delta},\ket{-\delta}}$.
+Altogether, the Hamiltonian is
 \begin{align}
   H = \sum_m \op{\bar m} \otimes
-  \sp{U \op{0} + \sum_{\delta>0} U_\delta X_\delta\p{U}}
+  \sp{U\op{0} + \sum_{\delta>0} U_\delta X_\delta\p{U}},
   &&
-  U_\delta \equiv \sqrt{U^2+\delta^2}.
+  U_\delta \equiv \sqrt{U^2+\delta^2},
 \end{align}
-The eigenvalues of $H$ are therefore $U$ and $\pm U_\delta$ for integer $\delta>0$.
-We now consider an initial state in which both large spins point at a polar angle $\theta$ and azimuthal angles $\pm\phi/2$:
+which has eigenvalues $U$ and $\pm U_\delta$ for integer $\delta>0$.
+
+\subsection{Initial state}
+
+We now consider an initial state with both large spins polarized at a polar angle $\theta$ and azimuthal angles $\pm\phi/2$:
 \begin{align}
@@
 
+\subsection{Observables}
+
 We now turn our attention to the measurement of collective observables.
@@
 where $\mu,\nu\in\set{-I,-I+1,\cdots,I}$ and $s_{\mu\nu j} = \op{\mu}{\nu}_j$ is a spin-transition operator for spin $j$.
-Rather than initial/final spin projections $\mu\to\nu$, it will be convenient to consider operators that induce spin transitions $\mu\to\mu+\epsilon$, and then expand these operators in the Dicke basis for spin-polarized states:
+Rather than initial/final spin projections $\nu\to\mu$, it will be convenient to consider operators that induce spin transitions $\mu\to\mu+\epsilon$, and then expand these operators in the spin-projection (Dicke) basis for spin-polarized states:
 \begin{align}
@@
 \begin{align}
-  S_{\mu\epsilon\sigma} \equiv \tilde S_{\mu\epsilon} \otimes \1
+  \tilde S_{\mu\epsilon}^\sigma \equiv \tilde S_{\mu\epsilon} \otimes \1
   + \sigma \1 \otimes \tilde S_{\mu\epsilon}
-  = \sum_{m,n} \xi_{\mu\epsilon}^m
-  \p{\op{m+\epsilon,n}{mn} + \sigma \op{n,m+\epsilon}{nm}},
+  = \sum_{m,n} \p{\xi_{\mu\epsilon}^m \ket{m+\epsilon,n}
+    + \sigma \xi_{\mu\epsilon}^n \ket{m,n+\epsilon}} \bra{mn},
 \end{align}
@@
 \begin{align}
-  S_{\mu\epsilon\sigma}= \sum_{m,\delta,\sigma} \p{\xi_{\mu\epsilon}^{m+\delta/2}
+  \tilde S_{\mu\epsilon}^\sigma
+  = \sum_{m,\delta,\sigma} \p{\xi_{\mu\epsilon}^{m+\delta/2}
     \ket{\bar m_\epsilon,\delta+\epsilon}

### 2020-10-29 sun_phases/spin_qudit_tomo.tex
added some references, and footnote explaining how to verify that 2d-1 measurement axes suffice

@@
 Quantum state tomography, or the task of reconstructing a quantum state by collecting and processing measurement data, is an essential primitive for quantum sensing, quantum simulation, and quantum information processing.
-The central importance of quantum state tomography has led to the development of a variety of techniques based on compressed sensing \cite{gross2010quantum}, matrix product states \cite{cramer2010efficient}, % confidence regions \cite{christandl2012reliable},
-maximum-likelihood estimation \cite{smolin2012efficient}, % linear \cite{qi2013quantum} and least-squares \cite{opatrny1997leastsquares} regression,
-Bayesian inference \cite{huszar2012adaptive}, and neural networks \cite{torlai2018neuralnetwork}, among others.
+The central importance of quantum state tomography has led to the development of a variety of techniques based on compressed sensing \cite{gross2010quantum}, matrix product states \cite{cramer2010efficient}, maximum-likelihood estimation \cite{smolin2012efficient}, Bayesian inference \cite{huszar2012adaptive}, and neural networks \cite{torlai2018neuralnetwork}, among others \cite{christandl2012reliable, qi2013quantum, opatrny1997leastsquares}.
 These techniques are typically developed in a general, information-theoretic setting, and make minimal assumptions about the physical medium of a quantum state.
@@
 Due in part to advancements in experimental capabilities to address nuclear spins in ultracold atomic systems [\red{CITE}], as well as developments in molecular cooling techniques [\red{CITE}], a particular setting of growing interest is the spin qudit, or a multilevel quantum angular momentum degree of freedom.
-Spin qudits can provide advantages over their qubit counterparts for quantum sensing \cite{evrard2019enhanced}, enable quantum simulations of SU($d$) magnetism \cite{banerjee2013atomic, cazalilla2014ultracold}, and offer unique capabilities for quantum error correction \cite{albert2020robust, gross2020encoding}.
+Spin qudits can provide advantages over their qubit counterparts for quantum sensing \cite{evrard2019enhanced}, enable quantum simulations of SU($d$) magnetism \cite{banerjee2013atomic, cazalilla2014ultracold, zhang2014spectroscopic, scazza2014observation, goban2018emergence}, and offer unique capabilities for quantum error correction \cite{albert2020robust, gross2020encoding}.
 In all cases, quantum state tomography is necessary to take full advantage of a spin qudit.
@@
 The problem of qudit tomography in particular is not new, with an extensive literature on a variety of techniques \cite{manko1997spin, thew2002qudit, flammia2005minimal, salazar2012quantum, sosa-martinez2017quantum, ha2018minimal, evrard2019enhanced, stefano2019set, palici2020oam}.
-However, existing proposals either require the capability to perform essentially arbitrary operations on a qudit \cite{thew2002qudit, flammia2005minimal, salazar2012quantum, sosa-martinez2017quantum, ha2018minimal, stefano2019set, palici2020oam}, or rely on infinite-dimensional representations of a quantum spin \cite{manko1997spin, evrard2019enhanced}, resulting in tomographic protocols that can be highly inefficient or even impossible in practice.
+However, existing proposals either require the capability to perform essentially arbitrary operations on a qudit \cite{thew2002qudit, flammia2005minimal, salazar2012quantum, sosa-martinez2017quantum, ha2018minimal, stefano2019set, palici2020oam}, or rely on infinite-dimensional representations of a quantum spin \cite{manko1997spin, evrard2019enhanced}, resulting in tomographic protocols that can be highly inefficient or unachievable in practice.
 In particular, the protocols based on infinite-dimensional representations of a quantum spin require only the capability to measure spin projection onto different spatial axes, which is usually straightforward with a spin qudit.
-Nonetheless, these protocols obfuscate the requirements for performing full state tomography, provide no straightforward error bounds for reconstructed states, and extract only a vanishing faction of the information contained in measurement data.
+Nonetheless, these protocols obfuscate the basic requirements for performing full state tomography, provide no straightforward error bounds or guarantees of accuracy, and extract only a vanishing faction of the information contained in measurement data.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Tomography as signal recovery on the sphere}
+\section{Spin tomography as signal recovery on the sphere}
 \label{sec:signal_recovery}
@@
 Unfortunately, we have no general proof of this conjecture, which may require a deeper analysis of fixed-degree signal recovery on the sphere.
-Nonetheless, we have used the tomography protocol in the next section to numerically verify that $2d-1$ measurement axes suffice for all $d\le200$.
+Nonetheless, we have used the tomography protocol in the next section to numerically verify that $2d-1$ measurement axes suffice for all $d\le200$ \footnote{Specifically, for all $d\le200$ we chose $2d-1$ measurement axes $V=\set{\v v}\subset\Omega_2$ by randomly sampling points on the sphere, and checked that the corresponding overlap matrix $G_V$ defined in Eq.~\eqref{eq:overlap_matrix} has $d^2$ nonzero eigenvalues, which is a necessary and sufficient condition to reconstruct an arbitrary $d$-level spin qudit state from measurements of spin projection onto the axes in $V$}.
 

### 2020-10-29 sun_phases/spin_qudit_tomo.tex
added text about choosing the number of measurement axes

@@
 \newcommand{\B}{\mathcal{B}}
-\newcommand{\N}{\mathcal{N}}
+\newcommand{\M}{\mathcal{M}}
 \renewcommand{\O}{\mathcal{O}}
@@
   We consider the task of performing quantum state tomography on a $d$-state spin qudit, using only measurements of spin projection onto different quantization axes.
-  By mapping this quantum tomography problem onto the classical problem of signal recovery on the sphere, we prove that full reconstruction of arbitrary mixed states is possible, and requires a minimum number of measurement axes, $r_d$, that is bounded by $2d-1\le r_d\le d^2$.
-  We conjecture that $r_d=2d-1$, which we have verified for all $d\le200$ using numerical experiments.
-  We then provide a parallelized algorithm with $O\p{r^3}$ runtime for certifying upper bounds on the error with which measurements of spin projection along $r$ quantization axes can be used to reconstruct an unknown quantum state.
+  By mapping this quantum tomography problem onto the classical problem of signal recovery on the sphere, we prove that full reconstruction of arbitrary mixed states is possible, and requires a minimum number of measurement axes, $r_d^{\t{min}}$, that is bounded by $2d-1\le r_d^{\t{min}}\le d^2$.
+  We conjecture that $r_d^{\t{min}}=2d-1$, which have verified for all $d\le200$ using numerical experiments.
+  We then provide a parallelized algorithm with $O\p{r^3}$ runtime for certifying an upper bound on the error with which measurements of spin projection onto $r$ quantization axes can reconstruct an unknown quantum state.
   Our algorithm motivates a simple randomized tomography protocol, which we use to study the trade-off between the number of measurement axes and state reconstruction error at a fixed total number of measurements [\red{note: this part is not yet complete}].
@@
 one might expect that $2d-1$ suitably chosen measurement axes would always suffice to perform full tomography of a $d$-level spin qudit.
-In the language of classical signal recovery, we would expect there to exist a set of $2\ell_{\t{max}}+1$ points on the sphere that enable the recovery of any degree-$\ell$ function with $\ell\le\ell_{\t{max}}$ (given prior knowledge of $\ell$) \footnote{We could further conjecture that if {\it any} set of points $V\subset\Omega_2$ enables the recovery of degree-$\ell_{\t{max}}$ functions on the sphere, then $V$ would allow recovery degree-$\ell$ functions with $\ell\le\ell_{\t{max}}$.
+In the language of classical signal recovery, we would expect there to exist a set of $2\ell_{\t{max}}+1$ points on the sphere that enable the recovery of any function with fixed degree $\ell\le\ell_{\t{max}}$ (given prior knowledge of $\ell$) \footnote{We could further conjecture that if {\it any} set of points $V\subset\Omega_2$ enables the recovery of degree-$\ell_{\t{max}}$ functions on the sphere, then $V$ would allow recovery degree-$\ell$ functions with $\ell\le\ell_{\t{max}}$.
 However, this condition is stronger than that necessary to guarantee full recovery of a $d$-level spin qudit state with $2d-1$ measurement axes.}.
 Unfortunately, we have no general proof of this conjecture, which may require a deeper analysis of fixed-degree signal recovery on the sphere.
-Nonetheless, we have used the tomography protocol in the next section to numerically verify that $2d-1$ measurement axes suffice for all $d\le200$ \footnote{Specifically, for all $d\le200$ we chose $2d-1$ measurement axes $V=\set{\v v}\subset\Omega_2$ by randomly sampling points on the sphere, and checked that the corresponding overlap matrix $G_V$ defined in Eq.~\eqref{eq:overlap_matrix} has $d^2$ nonzero eigenvalues, which is a necessary and sufficient condition to reconstruct an arbitrary $d$-level spin qudit state from measurements of spin projection onto the axes in $V$}.
+Nonetheless, the tomography protocol in the next section can be used to numerically verify that $2d-1$ measurement axes suffice for all $d\le200$.
 
@@
 Intuitively, a good choice of axes should minimize the error with which one can reconstruct an unknown quantum state from associated measurement data.
-If we can quantify this idea, then we can try to optimize over different choices of measurement axes to find a set that (approximately) minimizes the error in reconstructed states.
+If we can quantify this intuition, then we can try to optimize over different choices of measurement axes to find a set that (approximately) minimizes the error in reconstructed states.
 
@@
 where we neglect cross terms with different degrees $\ell$ because the corresponding inner products vanish, i.e.~$\obk{T_{\v v\ell,0}|T_{\v v'\ell',0}}=0$ for all $\ell\ne\ell'$.
-Eigenvectors of $G_V$ define linearly independent linear combinations of the operators in $\v T_V$.
-If the operators in $\v T_V$ span the entire space of operators on a $d$-level spin qudit, which is a necessary and sufficient condition for $V$ to allow for full state tomography, then $G_V$ must have exactly $d^2$ nonzero eigenvalues.
+If the operators in $\v T_V$ span the entire ($d^2$-dimensional) space of operators on a $d$-level spin qudit, which is a necessary and sufficient condition for $V$ to allow for full state tomography, then $G_V$ must have exactly $d^2$ nonzero eigenvalues.
 Indexing these eigenvalues $G^V_k$ and the corresponding (normalized) eigenvectors $\v x^V_k\equiv(x^V_{k,1},x^V_{k,2},\cdots)$ by an integer $k\in\set{1,2,\cdots,d^2}$, we can construct the orthonormal operators
@@
   = \sum_k \Bbk{\abs{\tilde\rho^V_k - \rho^V_k}^2}
-  \le \epsilon^2 \N_V^2,
+  \le \epsilon^2 \M_V^2,
   \label{eq:bound}
@@
 \begin{align}
-  \N_V^2 \equiv \sum_k \f1{G^V_k}
-  \label{eq:norm}
+  \M_V^2 \equiv \sum_k \f1{G^V_k}
+  \label{eq:scale}
 \end{align}
-is a norm determined by the set of measurement axes $V$.
-Note that the sum in Eq.~\eqref{eq:norm} is performed over the $d^2$ largest eigenvalues of $G_V$, and therefore may be divergent if $G_V$ has fewer than $d^2$ nonzero eigenvalues.
-Altogether, diagonalizing each of $d$ blocks of $G_V$ takes $O(\abs{V}^3)$ time, so the norm $\N_V$ can be computed in $O(d\abs{V}^3)$ serial or $O(\abs{V}^3)$ parallel runtime.
+is a scale determined by the set of measurement axes $V$.
+Note that the sum in Eq.~\eqref{eq:scale} is performed over the $d^2$ largest eigenvalues of $G_V$, and therefore may be divergent if $G_V$ has fewer than $d^2$ nonzero eigenvalues.
+Altogether, diagonalizing each of $d$ blocks of $G_V$ takes $O(\abs{V}^3)$ time, so the scale $\M_V$ can be computed in $O(d\abs{V}^3)$ serial or $O(\abs{V}^3)$ parallel runtime.
 
-The bound in Eq.~\eqref{eq:bound} motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and variationally optimize the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the error norm $\N_V$.
-If $\abs{V}$ is too large for variational optimization to be practical, one can instead generate many random sets of measurement axes, and then choose the set that minimizes $\N_V$.
+The bound in Eq.~\eqref{eq:bound} motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and variationally optimize the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the scale $\M_V$.
+If $\abs{V}$ is too large for variational optimization to be practical, one can instead generate many random sets of measurement axes, and then choose the set that minimizes $\M_V$.
 After choosing a set of measurement axes $V$, experimentally characterize the expectation values $\bk{T_j}_\rho$.
-If the experimental measurements $E_j$ of $\bk{T_j}_\rho$ are accurate to within an error $\lesssim\epsilon$, then the state $\tilde\rho$ defined in Eq.~\eqref{eq:reconstructed_state} is an approximation of the true state $\rho$ to within a 2-norm distance of $\lesssim\epsilon\N_V$.
+If the experimental measurements $E_j$ of $\bk{T_j}_\rho$ are accurate to within an error $\lesssim\epsilon$, then the state $\tilde\rho$ defined in Eq.~\eqref{eq:reconstructed_state} is an approximation of the true state $\rho$ to within a 2-norm distance of $\lesssim\epsilon\M_V$.
 Furthermore, if the empirical estimate $\tilde\rho$ has negative eigenvalues, its 2-norm distance from $\rho$ can be further reduced with maximum-likelihood corrections \cite{smolin2012efficient}.
@@
 
-As a last point, we mention that the cost function $\N_V$ can also evaluate the quality of measurement axes chosen with existing methods for sparse signal recovery on the sphere \cite{rauhut2011sparse, alem2012sparse}.
+As a last point, we mention that the cost function $\M_V$ can also evaluate the quality of measurement axes chosen with existing methods for sparse signal recovery on the sphere \cite{rauhut2011sparse, alem2012sparse}.
 In practice, we find that uniform random sampling outperforms these methods, in the sense that it generates low-cost sets of measurement axes with higher probability.
@@
 \label{sec:axes}
+
+We conjecture, and numerically verify for $d\le200$, that measurements of spin projection along $2d-1$ axes suffice to reconstruct an arbitrary state of a $d$-level spin qudit [\red{todo: make figure of $1/\M_V$ (minimized over, say, 100 random choices of measurement axes) as a function of the measurement axis number, for different values of $d$}].
+Even so, $2d-1$ measurement axes may not be the ``optimal'' number to use.
+Increasing the number of measurement axes generally decreases the error scale $\M_V$, but at the cost of having to estimate more observables.
+At a fixed total number of experimental measurements, having to estimate more observables means that fewer measurements are devoted to estimating any particular observable, which in turn increases the error $\epsilon$ on those estimates.
+This trade-off begs the question: how many measurement axes should be chosen to perform spin qudit tomography?
+
+The reconstruction error bound in Eq.~\eqref{eq:bound} provides a straightforward answer: the number of measurement axes should be chosen to minimize the product $\epsilon^2\M_V^2$, which bounds the expected error in the estimate $\tilde\rho$ of an unknown state $\rho$.
+If there are a total of $N$ spin projection measurements onto $r$ quantization axes, then $\sim N/r$ measurements are used to estimate spin projection onto any given axis.
+The variance in these estimates is then $\epsilon^2\sim 1/(N/r)=r/N$, or simply $\epsilon^2\sim r$ at a fixed total number of measurements (assuming all measurements are independent).
+Altogether, the number of measurement axes $r$ should be chosen in such a way as to minimize $r\M_V^2$.
 

### 2020-10-30 sun_phases/spin_qudit_tomo.tex
minor wording changes

@@
 The same orthogonality relations in Eqs.~\eqref{eq:ortho_full} and \eqref{eq:ortho_sphere} are inherited by the adjoint rotation matrix elements $\D_{mn}^\ell$ through the fact $\R\p{\v\omega}$ is the adjoint action of $R\p{\v\omega}$ (see Appendix \ref{sec:ortho}), and the functions $\set{\D_{m,0}^\ell}$ restricted to $\Omega_2$ are similarly another orthogonal basis for the space $L^2\p{\Omega_2}$ of square-integrable functions on the sphere.
-[\red{todo: make figures visualizing $Y_{\ell m}$ and $\D^\ell_{m,0}$ for comparison}]
 
@@
 
-In fact, the problem of finding suitable axes $V$ and coefficients $C_{\ell m}\p{\v v}$ to satisfy Eq.~\eqref{eq:tomo_recovery}, solving which would enable the recovery of arbitrary spin qudit states, can be mapped onto the well-studied problem of signal recovery on the sphere \cite{mcewen2011novel, mcewen2011sampling, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality} [\red{todo: make figure to visualize the signal recovery problem}].
+In fact, the problem of finding suitable axes $V$ and coefficients $C_{\ell m}\p{\v v}$ to satisfy Eq.~\eqref{eq:tomo_recovery}, solving which would enable the recovery of arbitrary spin qudit states, can be mapped onto the well-studied problem of signal recovery on the sphere \cite{mcewen2011novel, mcewen2011sampling, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
 The signal recovery problem can be stated as follows: given a square-integrable function $f\in L^2\p{\Omega_2}$ on the sphere, with the harmonic expansion
@@
 
-We conjecture, and numerically verify for $d\le200$, that measurements of spin projection along $2d-1$ axes suffice to reconstruct an arbitrary state of a $d$-level spin qudit [\red{todo: make figure of $1/\M_V$ (minimized over, say, 100 random choices of measurement axes) as a function of the measurement axis number, for different values of $d$}].
+We conjecture, and numerically verify for $d\le200$, that measurements of spin projection along $2d-1$ axes suffice to reconstruct an arbitrary state of a $d$-level spin qudit (see Figure [\red{REF}]).
 Even so, $2d-1$ measurement axes may not be the ``optimal'' number to use.
 Increasing the number of measurement axes generally decreases the error scale $\M_V$, but at the cost of having to estimate more observables.
-At a fixed total number of experimental measurements, having to estimate more observables means that fewer measurements are devoted to estimating any particular observable, which in turn increases the error $\epsilon$ on those estimates.
+If we fix the total number of measurements (or experimental trials), then having to estimate more observables means that fewer measurements are devoted to estimating any particular observable, which in turn increases the error $\epsilon$ on those estimates.
 This trade-off begs the question: how many measurement axes should be chosen to perform spin qudit tomography?
@@
 If there are a total of $N$ spin projection measurements onto $r$ quantization axes, then $\sim N/r$ measurements are used to estimate spin projection onto any given axis.
-The variance in these estimates is then $\epsilon^2\sim 1/(N/r)=r/N$, or simply $\epsilon^2\sim r$ at a fixed total number of measurements (assuming all measurements are independent).
-Altogether, the number of measurement axes $r$ should be chosen in such a way as to minimize $r\M_V^2$.
+The variance in estimates of spin projection is then $\epsilon^2\sim r/N$, or simply $\epsilon^2\sim r$ at a fixed total number of measurements $N$.
+Altogether, the number of measurement axes should be chosen in such a way as to minimize $r\M_V^2$ with $\abs{V}=r$.
 

### 2020-10-30 sun_phases/spin_qudit_tomo.tex
add discussion of computational complexity, place the tomography protocol in its own section

@@
   By mapping this quantum tomography problem onto the classical problem of signal recovery on the sphere, we prove that full reconstruction of arbitrary mixed states is possible, and requires a minimum number of measurement axes, $r_d^{\t{min}}$, that is bounded by $2d-1\le r_d^{\t{min}}\le d^2$.
-  We conjecture that $r_d^{\t{min}}=2d-1$, which have verified for all $d\le200$ using numerical experiments.
-  We then provide a parallelized algorithm with $O\p{r^3}$ runtime for certifying an upper bound on the error with which measurements of spin projection onto $r$ quantization axes can reconstruct an unknown quantum state.
+  We conjecture that $r_d^{\t{min}}=2d-1$, which we verify for all $d\le200$ using numerical experiments.
+  We then provide an algorithm with $O\p{d^4}$ runtime for certifying an upper bound on the error with which measurements of spin projection can reconstruct an unknown quantum state.
   Our algorithm motivates a simple randomized tomography protocol, which we use to study the trade-off between the number of measurement axes and state reconstruction error at a fixed total number of measurements [\red{note: this part is not yet complete}].
@@
 Unfortunately, we have no general proof of this conjecture, which may require a deeper analysis of fixed-degree signal recovery on the sphere.
-Nonetheless, the tomography protocol in the next section can be used to numerically verify that $2d-1$ measurement axes suffice for all $d\le200$.
+Nonetheless, we have numerically verified that $2d-1$ measurement axes suffice for all $d\le200$ (see Section \ref{sec:axes}).
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Reconstruction error bounds}
-\label{sec:protocol}
+\section{Reconstruction error bound}
+\label{sec:bound}
 
@@
 \end{align}
-Given empirical estimates $E_j$ of the expectation values $\bk{T_j}_\rho$, an empirical estimate $\tilde\rho$ of $\rho$ is then
+Given empirical estimates $E_j$ of the expectation values $\bk{T_j}_\rho$, an empirical estimate $\tilde\rho_V$ of $\rho$ is then
 \begin{align}
-  \tilde\rho \equiv \sum_k \tilde\rho^V_k Q^V_k,
+  \tilde\rho_V \equiv \sum_k \tilde\rho^V_k Q^V_k,
   \label{eq:reconstructed_state}
@@
 \end{align}
-The eigenvalues of the overlap matrix $G_V$ allow us to make concrete statements about the statistical error between the empirical estimate $\tilde\rho$ and the true state $\rho$.
+The eigenvalues of the overlap matrix $G_V$ allow us to make concrete statements about the statistical error between the empirical estimate $\tilde\rho_V$ and the true state $\rho$.
 Assume, for example, that the estimates $E_j$ are equal to $\bk{T_j}_\rho$ up to Gaussian noise with variance at most $\epsilon^2$, i.e.
@@
 \end{align}
-Using the fact that the operators $Q^V_k$ are orthonormal, we can therefore bound the mean squared 2-norm (Euclidean) distance between $\tilde\rho$ and $\rho$ as
+Using the fact that the operators $Q^V_k$ are orthonormal, we can therefore bound the mean squared 2-norm (Euclidean) distance between $\tilde\rho_V$ and $\rho$ as
 \begin{align}
-  \Bbk{\norm{\tilde\rho-\rho}_2^2}
+  \Bbk{\norm{\tilde\rho_V-\rho}_2^2}
   = \sum_k \Bbk{\abs{\tilde\rho^V_k - \rho^V_k}^2}
@@
 Note that the sum in Eq.~\eqref{eq:scale} is performed over the $d^2$ largest eigenvalues of $G_V$, and therefore may be divergent if $G_V$ has fewer than $d^2$ nonzero eigenvalues.
-Altogether, diagonalizing each of $d$ blocks of $G_V$ takes $O(\abs{V}^3)$ time, so the scale $\M_V$ can be computed in $O(d\abs{V}^3)$ serial or $O(\abs{V}^3)$ parallel runtime.
 
-The bound in Eq.~\eqref{eq:bound} motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and variationally optimize the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the scale $\M_V$.
+The complexity of computing $\M_V^2$ is determined by the need to:
+\begin{enumerate*}
+\item compute all rotated transition operators $T_{\v v\ell,0}$, which by the expansion $T_{\v v\ell,0}=R\p{\v v}T_{\ell,0}R\p{\v v}^\dag$ takes $O(\abs{V}d^4)$ serial or $O(d^3)$ parallel runtime;
+\item construct the overlap matrix $G_V$, whose block diagonal structure makes possible in $O(\abs{V}^2d^3)$ serial or $O(\abs{V}^2d^2)$ parallel runtime; and
+\item diagonalize the blocks of $G_V$, which takes $O(\abs{V}^3d)$ serial or $O(\abs{V}^3)$ parallel runtime.
+\end{enumerate*}
+Altogether, assuming $\abs{V}\propto d$ the bound in Eq.~\eqref{eq:bound} can be certified in $O(d^5)$ serial or $O(d^4)$ parallel runtime.
+In practice, for modest values of $d$ we find that vectorized operations can reduce serial runtime to $\sim O(r^4)$, and that serial execution time is divided about evenly between computing rotated transition operators $T_{\v v\ell,0}$ and constructing the overlap matrix $G_V$, with a negligible amount of time spent on diagonalization (see Figure [\red{REF}]).
+[\red{todo: add figure showing runtime for each routine as a function of $d$ with $\abs{V}=2d-1$; modify text appropriately after making the figure}].
+We suspect that more efficient representations of the transition operators $T_{\ell m}$, namely as $\p{2\ell+1}$-component vectors rather than $d\times d$ matrices, together with appropriate changes to the representation of the adjoint rotation operator $\R\p{\v v}$, should reduce the computational complexity of computing $\M_V^2$ by a factor of $O(d)$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Tomography protocols}
+\label{sec:protocol}
+
+The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and variationally optimize the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the error scale $\M_V$ defined in Eq.~\eqref{eq:scale}.
 If $\abs{V}$ is too large for variational optimization to be practical, one can instead generate many random sets of measurement axes, and then choose the set that minimizes $\M_V$.
 After choosing a set of measurement axes $V$, experimentally characterize the expectation values $\bk{T_j}_\rho$.
-If the experimental measurements $E_j$ of $\bk{T_j}_\rho$ are accurate to within an error $\lesssim\epsilon$, then the state $\tilde\rho$ defined in Eq.~\eqref{eq:reconstructed_state} is an approximation of the true state $\rho$ to within a 2-norm distance of $\lesssim\epsilon\M_V$.
-Furthermore, if the empirical estimate $\tilde\rho$ has negative eigenvalues, its 2-norm distance from $\rho$ can be further reduced with maximum-likelihood corrections \cite{smolin2012efficient}.
-In terms of classical computing costs before and after estimating expectation values $\bk{T_j}_\rho$, this tomography protocol is bottlenecked by the need to diagonalize the overlap matrix $G_V$, which has $d$ blocks of size $\abs{V}\times\abs{V}$.
-Altogether, this protocol therefore has an $O(d\abs{V}^3)$ serial or $O(\abs{V}^3)$ parallel runtime.
+If the experimental measurements $E_j$ of $\bk{T_j}_\rho$ are accurate to within an error $\lesssim\epsilon$, then the state $\tilde\rho_V$ defined in Eq.~\eqref{eq:reconstructed_state} is an approximation of the true state $\rho$ to within a 2-norm distance $\lesssim\epsilon\M_V$.
+Furthermore, if the empirical estimate $\tilde\rho_V$ has negative eigenvalues, its 2-norm distance from $\rho$ can be further reduced with maximum-likelihood corrections \cite{smolin2012efficient}.
 
@@
 
-As a last point, we mention that the cost function $\M_V$ can also evaluate the quality of measurement axes chosen with existing methods for sparse signal recovery on the sphere \cite{rauhut2011sparse, alem2012sparse}.
-In practice, we find that uniform random sampling outperforms these methods, in the sense that it generates low-cost sets of measurement axes with higher probability.
+As a last point, we mention that the error scale $\M_V$ can also evaluate the ``quality'' of measurement axes $V$ chosen with existing methods for sparse signal recovery on the sphere \cite{rauhut2011sparse, alem2012sparse}.
+In practice, we find that uniform random sampling outperforms these methods, in the sense that it generates low-error sets of measurement axes with higher probability.
 Two observations may explain the poor performance of these sparse signal recovery methods.
-First, spin qudit tomography essentially boils down to recovering sparse signals with a particular harmonic structure, namely with a fixed degree $\ell$.
-The existing sparse signal recovery algorithms do not assume or account for any harmonic structure, and may therefore perform worse than expected on this particular subset of sparse signals.
+First, spin qudit tomography essentially boils down to recovering sparse signals with a particular harmonic structure, namely a fixed degree $\ell$.
+The existing sparse signal recovery algorithms do not assume or account for any harmonic structure, and may therefore perform worse than expected on this particular subset of signals.
 Second, spin qudit tomography requires recovering signals that are band-limited at degree $d$, but have $O(d)$ non-zero harmonic coefficients (out of $d^2$ coefficients in total).
@@
 
-The reconstruction error bound in Eq.~\eqref{eq:bound} provides a straightforward answer: the number of measurement axes should be chosen to minimize the product $\epsilon^2\M_V^2$, which bounds the expected error in the estimate $\tilde\rho$ of an unknown state $\rho$.
+The reconstruction error bound in Eq.~\eqref{eq:bound} provides a straightforward answer: the number of measurement axes should be chosen to minimize the product $\epsilon^2\M_V^2$, which bounds the expected error in the estimate $\tilde\rho_V$ of an unknown state $\rho$.
 If there are a total of $N$ spin projection measurements onto $r$ quantization axes, then $\sim N/r$ measurements are used to estimate spin projection onto any given axis.

### 2020-11-02 sun_phases/spin_qudit_tomo.tex
minor claims to commentary about computing the overlap matrix, and added a reference to a yet unwritten appendix about computing the error bound

@@
 \end{enumerate*}
+We discuss these calculations in more detail in Appendix \ref{sec:overlaps}.
 Altogether, assuming $\abs{V}\propto d$ the bound in Eq.~\eqref{eq:bound} can be certified in $O(d^5)$ serial or $O(d^4)$ parallel runtime.
-In practice, for modest values of $d$ we find that vectorized operations can reduce serial runtime to $\sim O(r^4)$, and that serial execution time is divided about evenly between computing rotated transition operators $T_{\v v\ell,0}$ and constructing the overlap matrix $G_V$, with a negligible amount of time spent on diagonalization (see Figure [\red{REF}]).
+In practice, for modest values of $d$ we find that serial execution time is divided about evenly between computing rotated transition operators $T_{\v v\ell,0}$ and constructing the overlap matrix $G_V$, with a negligible amount of time spent on diagonalization (see Figure [\red{REF}]).
 [\red{todo: add figure showing runtime for each routine as a function of $d$ with $\abs{V}=2d-1$; modify text appropriately after making the figure}].
-We suspect that more efficient representations of the transition operators $T_{\ell m}$, namely as $\p{2\ell+1}$-component vectors rather than $d\times d$ matrices, together with appropriate changes to the representation of the adjoint rotation operator $\R\p{\v v}$, should reduce the computational complexity of computing $\M_V^2$ by a factor of $O(d)$.
+We suspect that more efficient representations of the transition operators $T_{\ell m}$, namely as $\p{2\ell+1}$-component vectors rather than $d\times d$ matrices, together with appropriate changes to the representation of the adjoint rotation operator $\R\p{\v v}$, can reduce the complexity of computing $\M_V^2$ by a factor of $O(d)$.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Computing the overlap matrix}
+\label{sec:overlaps}
+
+[\red{todo: walk through the calculation of the error scale $\M_V$.}]
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Transition operator product expansion}

### 2020-11-02 sun_phases/spin_qudit_tomo.tex
major changes to the section on the reconstruction error bound, which now constructs a 'measurement matrix' and computes its singular values (rather than constructing an 'overlap matrix' and computing its eigenvalues)

--- 2020-10-30 sun_phases/spin_qudit_tomo.tex
@@
 \newcommand{\B}{\mathcal{B}}
-\newcommand{\M}{\mathcal{M}}
 \renewcommand{\O}{\mathcal{O}}
@@
 \newcommand{\R}{\mathcal{R}}
+\renewcommand{\S}{\mathcal{S}}
 \newcommand{\T}{\mathcal{T}}
@@
 
-A set of measurement axes $V=\set{\v v}$ nominally induces a set of projectors $\set{\Pi_{\v v\mu}^V}$ that will be measured in an experiment, where for clarify we index objects by $V$ to emphasize their dependence on the choice of measurement axes.
+A set of measurement axes $V=\set{\v v}$ nominally induces a set of projectors $\set{\Pi_{\v v\mu}}$ that will be measured in an experiment.
 By a simple change of basis, measuring these projectors is equivalent to measuring the set of transition operators $\v T_V\equiv\set{T_{\v v\ell,0}}$.
-The operators in $\v T_V$ can be used to construct the overlap (Gram) matrix
+By flattening qudit operators $X = \sum_{\mu,\nu} X_{\mu\nu} \op{\mu}{\nu}$ into the $d^2$-component vectors $\oket{X} \equiv \sum_{\mu,\nu} X_{\mu\nu} \ket{\mu\nu}$, we can construct the {\it measurement matrix}
 \begin{align}
-  G_V \equiv \sum_{\v v,\v v',\ell}
-  \obk{T_{\v v\ell,0}|T_{\v v'\ell,0}} \op{\v v\ell}{\v v'\ell},
-  \label{eq:overlap_matrix}
+  M_V \equiv \sum_{\v v,\ell} \oket{T_{\v v\ell,0}} \bra{\v v\ell},
+  \label{eq:meas_mat}
 \end{align}
-where we neglect cross terms with different degrees $\ell$ because the corresponding inner products vanish, i.e.~$\obk{T_{\v v\ell,0}|T_{\v v'\ell',0}}=0$ for all $\ell\ne\ell'$.
-If the operators in $\v T_V$ span the entire ($d^2$-dimensional) space of operators on a $d$-level spin qudit, which is a necessary and sufficient condition for $V$ to allow for full state tomography, then $G_V$ must have exactly $d^2$ nonzero eigenvalues.
-Indexing these eigenvalues $G^V_k$ and the corresponding (normalized) eigenvectors $\v x^V_k\equiv(x^V_{k,1},x^V_{k,2},\cdots)$ by an integer $k\in\set{1,2,\cdots,d^2}$, we can construct the orthonormal operators
+whose columns are flattenings of the operators in $\v T_V$.
+A necessary and sufficient for $V$ to allow for full state tomography is for the operators in $\v T_V$, or equivalently the columns of $M_V$, to span the entire ($d^2$-dimensional) space of operators on a $d$-level spin qudit.
+In this case, $M_V$ must be full rank, with $d^2$ nonzero singular values.
+Indexing these singular values $M^V_k$ and the corresponding (normalized) right singular vectors $\v x^V_k\equiv(x^V_{k,1},x^V_{k,2},\cdots)$ by an integer $k\in\set{1,2,\cdots,d^2}$, we can construct the orthonormal qudit operators
 \begin{align}
-  Q^V_k \equiv \sum_j g^V_{kj} T_j,
+  Q^V_k \equiv \sum_j q^V_{kj} T_j,
   &&
-  g^V_{kj} \equiv \f{x^V_{kj}}{\sqrt{G^V_k}},
+  q^V_{kj} \equiv \f{x^V_{kj}}{M^V_k},
 \end{align}
@@
 \begin{align}
-  \tilde\rho^V_k \equiv \sum_j g^V_{kj} E_j
-  \approx \sum_j g^V_{kj} \bk{T_j}_\rho
+  \tilde\rho^V_k \equiv \sum_j q^V_{kj} E_j
+  \approx \sum_j q^V_{kj} \bk{T_j}_\rho
   = \bk{Q^V_k}_\rho
@@
 \end{align}
-The eigenvalues of the overlap matrix $G_V$ allow us to make concrete statements about the statistical error between the empirical estimate $\tilde\rho_V$ and the true state $\rho$.
+The singular values of $M_V$ allow us to make concrete statements about the statistical error between the empirical estimate $\tilde\rho_V$ and the true state $\rho$.
 Assume, for example, that the estimates $E_j$ are equal to $\bk{T_j}_\rho$ up to Gaussian noise with variance at most $\epsilon^2$, i.e.
@@
   &= \Bbk{\p{\tilde\rho^V_k-\rho^V_k}^*\p{\tilde\rho^V_k-\rho^V_k}} \\
-  &= \sum_{j,j'} \p{g^V_{kj}}^* g^V_{kj'}\,
+  &= \sum_{j,j'} \p{q^V_{kj}}^* q^V_{kj'}\,
   \bbk{\epsilon_j \epsilon_{j'}} \\
-  &\le \sum_j \abs{g^V_{kj}}^2 \epsilon^2
-  = \f{\epsilon^2}{G^V_k}.
+  &\le \sum_j \abs{q^V_{kj}}^2 \epsilon^2
+  = \p{\f{\epsilon}{M^V_k}}^2.
 \end{align}
@@
   = \sum_k \Bbk{\abs{\tilde\rho^V_k - \rho^V_k}^2}
-  \le \epsilon^2 \M_V^2,
+  \le \epsilon^2 \S_V^2,
   \label{eq:bound}
@@
 \begin{align}
-  \M_V^2 \equiv \sum_k \f1{G^V_k}
+  \S_V \equiv \norm{M_V^{-1}}_2 = \sqrt{\sum_k\f1{\p{M^V_k}^2}},
   \label{eq:scale}
 \end{align}
-is a scale determined by the set of measurement axes $V$.
-Note that the sum in Eq.~\eqref{eq:scale} is performed over the $d^2$ largest eigenvalues of $G_V$, and therefore may be divergent if $G_V$ has fewer than $d^2$ nonzero eigenvalues.
+which we call the {\it error scale} of $V$, is the 2-norm of the pseudoinverse $M_V^{-1}$ of $M_V$.
+Note that the bound in Eq.~\eqref{eq:bound} only holds if the measurement matrix $M_V$ is nonsingular, as otherwise some of the operators $Q^V_k$ that we used to expand $\rho$ and $\tilde\rho_V$ are undefined, reflecting the fact that if $M_V$ is singular then measuring spin projections along all axes in $V$ does not provide sufficient information to reconstruct arbitrary quantum states.
 
-The complexity of computing $\M_V^2$ is determined by the need to:
+The complexity of computing the error scale $\S_V$ is nominally determined by the need to
 \begin{enumerate*}
-\item compute all rotated transition operators $T_{\v v\ell,0}$, which by the expansion $T_{\v v\ell,0}=R\p{\v v}T_{\ell,0}R\p{\v v}^\dag$ takes $O(\abs{V}d^4)$ serial or $O(d^3)$ parallel runtime;
-\item construct the overlap matrix $G_V$, whose block diagonal structure makes possible in $O(\abs{V}^2d^3)$ serial or $O(\abs{V}^2d^2)$ parallel runtime; and
-\item diagonalize the blocks of $G_V$, which takes $O(\abs{V}^3d)$ serial or $O(\abs{V}^3)$ parallel runtime.
+\item construct the measurement matrix $M_V$, and
+\item compute the singular values of $M_V$.
 \end{enumerate*}
-Altogether, assuming $\abs{V}\propto d$ the bound in Eq.~\eqref{eq:bound} can be certified in $O(d^5)$ serial or $O(d^4)$ parallel runtime.
-In practice, for modest values of $d$ we find that vectorized operations can reduce serial runtime to $\sim O(r^4)$, and that serial execution time is divided about evenly between computing rotated transition operators $T_{\v v\ell,0}$ and constructing the overlap matrix $G_V$, with a negligible amount of time spent on diagonalization (see Figure [\red{REF}]).
+By the expansion $T_{\v v\ell,0} = R\p{\v v} T_{\ell,0} R\p{\v v}^\dag$, each of the $\abs{V}d$ columns of $M_V$ takes $O(d^3)$ time to compute, for an overall $O(\abs{V}d^4)$ serial or $O(d^3)$ parallel runtime.
+In order to compute the singular values of $M_V$, we first note that the orthogonality of transition operators with different degrees, i.e.~$\obk{T_{\v v\ell,0}|T_{\v v',\ell',0}} = 0$ for $\ell\ne\ell'$, implies that $M_V$ can be decomposed into a direct sum of the rectangular blocks
+\begin{align}
+  M_{V\ell} \equiv \sum_{\v v} \oket{T_{\v v\ell,0}} \bra{\v v},
+\end{align}
+using which we can expand
+\begin{align}
+  \S_V^2 = \sum_\ell \S_{V\ell}^2,
+  &&
+  \S_{V\ell} \equiv \norm{M_{V\ell}^{-1}}_2,
+\end{align}
+where $M_{V\ell}^{-1}$ is the pseudoinverse of $M_{V\ell}$.
+Finding the singular values of each block $M_{V\ell}$ takes $O(\abs{V}^2d^2)$ time at most \footnote{The calculation of $\S_{V\ell}$ can be sped up by the fact that it is independent of $d$ as long as $d\ge\ell+1$, so we can essentially set $d=\ell+1$ for the purposes of computing $\S_{V\ell}$}.
+Altogether, if we assume that $\abs{V}\sim d$ then certifying the bound in Eq.~\eqref{eq:bound} by computing $\S_V$ takes $O(d^5)$ serial or $O(d^4)$ parallel runtime (see Figure [\red{REF}]).
 [\red{todo: add figure showing runtime for each routine as a function of $d$ with $\abs{V}=2d-1$; modify text appropriately after making the figure}].
-We suspect that more efficient representations of the transition operators $T_{\ell m}$, namely as $\p{2\ell+1}$-component vectors rather than $d\times d$ matrices, together with appropriate changes to the representation of the adjoint rotation operator $\R\p{\v v}$, should reduce the computational complexity of computing $\M_V^2$ by a factor of $O(d)$.
+We suspect that more efficient representations of the transition operators $T_{\ell m}$, namely as $\p{2\ell+1}$-component vectors rather than $d\times d$ matrices, together with appropriate changes to the representation of the adjoint rotation operator $\R\p{\v v}$, can reduce the complexity of computing $\S_V$ by a factor of $O(d)$.
 
@@
 
-The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and variationally optimize the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the error scale $\M_V$ defined in Eq.~\eqref{eq:scale}.
-If $\abs{V}$ is too large for variational optimization to be practical, one can instead generate many random sets of measurement axes, and then choose the set that minimizes $\M_V$.
+The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and variationally optimize the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the error scale $\S_V$ defined in Eq.~\eqref{eq:scale}.
+If $\abs{V}$ is too large for variational optimization to be practical, one can instead generate many random sets of measurement axes, and then choose the set that minimizes $\S_V$.
 After choosing a set of measurement axes $V$, experimentally characterize the expectation values $\bk{T_j}_\rho$.
-If the experimental measurements $E_j$ of $\bk{T_j}_\rho$ are accurate to within an error $\lesssim\epsilon$, then the state $\tilde\rho_V$ defined in Eq.~\eqref{eq:reconstructed_state} is an approximation of the true state $\rho$ to within a 2-norm distance $\lesssim\epsilon\M_V$.
+If the experimental measurements $E_j$ of $\bk{T_j}_\rho$ are accurate to within an error $\lesssim\epsilon$, then the state $\tilde\rho_V$ defined in Eq.~\eqref{eq:reconstructed_state} is an approximation of the true state $\rho$ to within a 2-norm distance $\lesssim\epsilon\S_V$.
 Furthermore, if the empirical estimate $\tilde\rho_V$ has negative eigenvalues, its 2-norm distance from $\rho$ can be further reduced with maximum-likelihood corrections \cite{smolin2012efficient}.
@@
 
-As a last point, we mention that the error scale $\M_V$ can also evaluate the ``quality'' of measurement axes $V$ chosen with existing methods for sparse signal recovery on the sphere \cite{rauhut2011sparse, alem2012sparse}.
-In practice, we find that uniform random sampling outperforms these methods, in the sense that it generates low-error sets of measurement axes with higher probability.
+As a last point, we mention that the error scale $\S_V$ can also evaluate the ``quality'' of measurement axes $V$ chosen with existing methods for sparse signal recovery on the sphere \cite{rauhut2011sparse, alem2012sparse}.
+In practice, we find that uniform random sampling outperforms these methods, in the sense that it generates low-error-scale sets of measurement axes with higher probability.
 Two observations may explain the poor performance of these sparse signal recovery methods.
@@
 Even so, $2d-1$ measurement axes may not be the ``optimal'' number to use.
-Increasing the number of measurement axes generally decreases the error scale $\M_V$, but at the cost of having to estimate more observables.
+Increasing the number of measurement axes generally decreases the error scale $\S_V$, but at the cost of having to estimate more observables.
 If we fix the total number of measurements (or experimental trials), then having to estimate more observables means that fewer measurements are devoted to estimating any particular observable, which in turn increases the error $\epsilon$ on those estimates.
@@
 
-The reconstruction error bound in Eq.~\eqref{eq:bound} provides a straightforward answer: the number of measurement axes should be chosen to minimize the product $\epsilon^2\M_V^2$, which bounds the expected error in the estimate $\tilde\rho_V$ of an unknown state $\rho$.
+The reconstruction error bound in Eq.~\eqref{eq:bound} provides a straightforward answer: the number of measurement axes should be chosen to minimize the product $\epsilon^2\S_V^2$, which bounds the expected error in the estimate $\tilde\rho_V$ of an unknown state $\rho$.
 If there are a total of $N$ spin projection measurements onto $r$ quantization axes, then $\sim N/r$ measurements are used to estimate spin projection onto any given axis.
 The variance in estimates of spin projection is then $\epsilon^2\sim r/N$, or simply $\epsilon^2\sim r$ at a fixed total number of measurements $N$.
-Altogether, the number of measurement axes should be chosen in such a way as to minimize $r\M_V^2$ with $\abs{V}=r$.
+Altogether, the number of measurement axes should be chosen in such a way as to minimize $r\S_V^2$ with $\abs{V}=r$.
 

### 2020-11-02 sun_phases/spin_qudit_tomo.tex
further reduce calculation of singular values

@@
 \email{mika.perlin@gmail.com}
+% \author{Diego Barberena}
 \author{Ana Maria Rey}
@@
 
-A set of measurement axes $V=\set{\v v}$ nominally induces a set of projectors $\set{\Pi_{\v v\mu}}$ that will be measured in an experiment.
-By a simple change of basis, measuring these projectors is equivalent to measuring the set of transition operators $\v T_V\equiv\set{T_{\v v\ell,0}}$.
+A choice of measurement axes $V=\set{\v v}$ determines a set of projectors $\v\Pi_V \equiv \set{\Pi_{\v v\mu}}$ that will be measured in an experiment.
 By flattening qudit operators $X = \sum_{\mu,\nu} X_{\mu\nu} \op{\mu}{\nu}$ into the $d^2$-component vectors $\oket{X} \equiv \sum_{\mu,\nu} X_{\mu\nu} \ket{\mu\nu}$, we can construct the {\it measurement matrix}
 \begin{align}
-  M_V \equiv \sum_{\v v,\ell} \oket{T_{\v v\ell,0}} \bra{\v v\ell},
+  M_V \equiv \sum_{\v v,\ell} \oket{\Pi_{\v v\mu}} \bra{\v v\mu},
   \label{eq:meas_mat}
 \end{align}
-whose columns are flattenings of the operators in $\v T_V$.
-A necessary and sufficient for $V$ to allow for full state tomography is for the operators in $\v T_V$, or equivalently the columns of $M_V$, to span the entire ($d^2$-dimensional) space of operators on a $d$-level spin qudit.
-In this case, $M_V$ must be full rank, with $d^2$ nonzero singular values.
+whose columns are flattenings of the operators in $\v\Pi_V$.
+A necessary and sufficient condition for $V$ to allow for full state tomography is that the projectors in $\v\Pi_V$, or equivalently the columns of $M_V$, to span the entire ($d^2$-dimensional) space of operators on a $d$-level spin qudit.
+In this case $M_V$ must be full rank, with $d^2$ nonzero singular values.
 Indexing these singular values $M^V_k$ and the corresponding (normalized) right singular vectors $\v x^V_k\equiv(x^V_{k,1},x^V_{k,2},\cdots)$ by an integer $k\in\set{1,2,\cdots,d^2}$, we can construct the orthonormal qudit operators
 \begin{align}
-  Q^V_k \equiv \sum_j q^V_{kj} T_j,
+  Q^V_k \equiv \sum_j q^V_{kj} \Pi_j,
   &&
-  q^V_{kj} \equiv \f{x^V_{kj}}{M^V_k},
+  \v q^V_k \equiv \f{\v x^V_k}{M^V_k},
 \end{align}
-where for shorthand we use a combined index $j=\p{\v v_j,\ell_j}$ to specify both a measurement axis $\v v_j$ and a degree $\ell_j$, which identify the transition operator $T_j\equiv T_{\v v_j\ell_j,0}$.
+where for shorthand we use a combined index $j=\p{\v v_j,\mu_j}$ to specify both a measurement axis $\v v_j$ and a spin projection $\mu_j$, which identify the projector operator $\Pi_j\equiv \Pi_{\v v_j\mu_j}$.
 These operators allow us to expand any state $\rho$ of a $d$-level spin qudit in the form
@@
 \end{align}
-Given empirical estimates $E_j$ of the expectation values $\bk{T_j}_\rho$, an empirical estimate $\tilde\rho_V$ of $\rho$ is then
+Given empirical estimates $E_j$ of the expectation values $\bk{\Pi_j}_\rho$, an empirical estimate $\tilde\rho_V$ of $\rho$ is then
 \begin{align}
@@
   \tilde\rho^V_k \equiv \sum_j q^V_{kj} E_j
-  \approx \sum_j q^V_{kj} \bk{T_j}_\rho
+  \approx \sum_j q^V_{kj} \bk{\Pi_j}_\rho
   = \bk{Q^V_k}_\rho
@@
 The singular values of $M_V$ allow us to make concrete statements about the statistical error between the empirical estimate $\tilde\rho_V$ and the true state $\rho$.
-Assume, for example, that the estimates $E_j$ are equal to $\bk{T_j}_\rho$ up to Gaussian noise with variance at most $\epsilon^2$, i.e.
+Assume, for example, that the estimates $E_j$ are equal to $\bk{\Pi_j}_\rho$ up to uncorrelated noise with variance at most $\epsilon^2$, i.e.
 \begin{align}
-  E_j = \bk{T_j}_\rho + \epsilon_j,
+  E_j = \bk{\Pi_j}_\rho + \epsilon_j,
   &&
@@
 \end{align}
-where $\set{\epsilon_j}$ are independent random variables, and we use the double brackets $\bbk{\cdot}$ to denote statistical averaging.
+where $\set{\epsilon_j}$ are independent random variables and we use the double brackets $\bbk{\cdot}$ to denote statistical averaging.
 The mean squared error with which $\tilde\rho^V_k$ approximates $\rho^V_k$ is then
@@
 \end{align}
-where
+where the {\it error scale} $\S_V$ of $V$, defined by
 \begin{align}
@@
 \end{align}
-which we call the {\it error scale} of $V$, is the 2-norm of the pseudoinverse $M_V^{-1}$ of $M_V$.
+is the 2-norm of the pseudoinverse $M_V^{-1}$ of $M_V$.
 Note that the bound in Eq.~\eqref{eq:bound} only holds if the measurement matrix $M_V$ is nonsingular, as otherwise some of the operators $Q^V_k$ that we used to expand $\rho$ and $\tilde\rho_V$ are undefined, reflecting the fact that if $M_V$ is singular then measuring spin projections along all axes in $V$ does not provide sufficient information to reconstruct arbitrary quantum states.
 
-The complexity of computing the error scale $\S_V$ is nominally determined by the need to
+The complexity of computing the error scale $\S_V$ is determined by the need to
 \begin{enumerate*}
@@
 \end{enumerate*}
-By the expansion $T_{\v v\ell,0} = R\p{\v v} T_{\ell,0} R\p{\v v}^\dag$, each of the $\abs{V}d$ columns of $M_V$ takes $O(d^3)$ time to compute, for an overall $O(\abs{V}d^4)$ serial or $O(d^3)$ parallel runtime.
-In order to compute the singular values of $M_V$, we first note that the orthogonality of transition operators with different degrees, i.e.~$\obk{T_{\v v\ell,0}|T_{\v v',\ell',0}} = 0$ for $\ell\ne\ell'$, implies that $M_V$ can be decomposed into a direct sum of the rectangular blocks
+These tasks can be drastically simplified by use of the transition operators.
+Specifically, the orthogonality of transition operators with different degrees, i.e.~$\obk{T_{\v v\ell,0}|T_{\v v',\ell',0}} = 0$ for $\ell\ne\ell'$, implies that $M_V$ has a block-diagonal structure in the transition operator basis, with rectangular blocks
 \begin{align}
-  M_{V\ell} \equiv \sum_{\v v} \oket{T_{\v v\ell,0}} \bra{\v v},
+  M_{V\ell} \equiv U_\ell \sum_{\v v} \oket{T_{\v v\ell,0}} \bra{\v v},
+  &&
+  U_\ell \equiv \sum_{m=-\ell}^\ell \ket{m} \obra{T_{\ell m}},
 \end{align}
-using which we can expand
+that we can use to expand
 \begin{align}
@@
 where $M_{V\ell}^{-1}$ is the pseudoinverse of $M_{V\ell}$.
-Finding the singular values of each block $M_{V\ell}$ takes $O(\abs{V}^2d^2)$ time at most \footnote{The calculation of $\S_{V\ell}$ can be sped up by the fact that it is independent of $d$ as long as $d\ge\ell+1$, so we can essentially set $d=\ell+1$ for the purposes of computing $\S_{V\ell}$}.
-Altogether, if we assume that $\abs{V}\sim d$ then certifying the bound in Eq.~\eqref{eq:bound} by computing $\S_V$ takes $O(d^5)$ serial or $O(d^4)$ parallel runtime (see Figure [\red{REF}]).
-[\red{todo: add figure showing runtime for each routine as a function of $d$ with $\abs{V}=2d-1$; modify text appropriately after making the figure}].
-We suspect that more efficient representations of the transition operators $T_{\ell m}$, namely as $\p{2\ell+1}$-component vectors rather than $d\times d$ matrices, together with appropriate changes to the representation of the adjoint rotation operator $\R\p{\v v}$, can reduce the complexity of computing $\S_V$ by a factor of $O(d)$.
+By the expansion $T_{\v v\ell,0} = R\p{\v v} T_{\ell,0} R\p{\v v}^\dag$, the $\p{2\ell+1}\times\abs{V}$ matrix $M_{V\ell}$ can be constructed in $O(\abs{V}d^3)$ time, and computing its singular values takes at most $O(\abs{V}d^2)$ time \footnote{The calculation of $\S_{V\ell}$ can be sped up by the fact that its value is independent of $d$ (as long as $d\ge\ell+1$), so we can essentially set $d=\ell+1$ for the purposes of computing $\S_{V\ell}$}.
+Altogether, if we assume that $\abs{V}\sim d$ then computing $\S_V$ to certify the bound in Eq.~\eqref{eq:bound} takes $O(d^5)$ serial or $O(d^4)$ parallel runtime (see Figure [\red{REF}]).
+[\red{todo: add figure showing runtimes as a function of $d$ with $\abs{V}=2d-1$; modify text appropriately after making the figure}].
+
+[\red{todo: add a careful derivation of the error bound that assumes $n$ measurements of spin projection along each axis into an Appendix.  Briefly discuss the result of this derivation here.}]
 
@@
 If $\abs{V}$ is too large for variational optimization to be practical, one can instead generate many random sets of measurement axes, and then choose the set that minimizes $\S_V$.
-After choosing a set of measurement axes $V$, experimentally characterize the expectation values $\bk{T_j}_\rho$.
-If the experimental measurements $E_j$ of $\bk{T_j}_\rho$ are accurate to within an error $\lesssim\epsilon$, then the state $\tilde\rho_V$ defined in Eq.~\eqref{eq:reconstructed_state} is an approximation of the true state $\rho$ to within a 2-norm distance $\lesssim\epsilon\S_V$.
+After choosing a set of measurement axes $V$, experimentally characterize the expectation values $\bk{\Pi_j}_\rho$.
+If the experimental measurements $E_j$ of $\bk{\Pi_j}_\rho$ are accurate to within an error $\lesssim\epsilon$, then the state $\tilde\rho_V$ defined in Eq.~\eqref{eq:reconstructed_state} is an approximation of the true state $\rho$ to within a 2-norm distance $\lesssim\epsilon\S_V$.
 Furthermore, if the empirical estimate $\tilde\rho_V$ has negative eigenvalues, its 2-norm distance from $\rho$ can be further reduced with maximum-likelihood corrections \cite{smolin2012efficient}.

### 2020-11-02 sun_phases/scripts/find_axes.py

@@ -28,6 +28,14 @@ axis_num = 2*dim-1
 
 diag_trans_ops = [ np.diag(transition_op(LL+1,LL,0)) for LL in range(dim) ]
 
+def unit_vector(dim, idx):
+    vec = np.zeros(dim)
+    vec[idx] = 1
+    return vec
+def proj_to_trans_degree(LL):
+    return np.vstack([ transition_op(LL+1,LL,MM).ravel() for MM in range(-LL,LL+1) ])
+proj_to_trans = [ proj_to_trans_degree(LL) for LL in range(dim) ]
+
 zhat = [1,0,0]
 xhat = [0,1,0]
 yhat = [0,0,1]
@@ -72,7 +80,8 @@ def axis_trans_op(degree, point):
 
 # compute all rotated (order-0) transition operators of a given degree
 def axes_trans_ops(degree, points):
-    return np.array([ axis_trans_op(degree, point).ravel() for point in points ])
+    proj_op_mat = np.array([ axis_trans_op(degree, point).ravel() for point in points ])
+    return proj_to_trans[degree] @ proj_op_mat.T
 
 # get squared singular values of a matrix, in decreasing order
 def get_sqr_svdvals(matrix):

### 2020-11-03 sun_phases/spin_qudit_tomo.tex
better explain the computation of the error scale

@@
 The singular values of $M_V$ allow us to make concrete statements about the statistical error between the empirical estimate $\tilde\rho_V$ and the true state $\rho$.
-Assume, for example, that the estimates $E_j$ are equal to $\bk{\Pi_j}_\rho$ up to uncorrelated noise with variance at most $\epsilon^2$, i.e.
+Assume, for example, that the estimates $E_j$ are equal to $\bk{\Pi_j}_\rho$ up to uncorrelated noise with variance at most $\epsilon^2$:
 \begin{align}
@@
 \end{align}
-where $\set{\epsilon_j}$ are independent random variables and we use the double brackets $\bbk{\cdot}$ to denote statistical averaging.
+where $\set{\epsilon_j}$ are independent random variables and we use the double brackets $\bbk{\cdot}$ to denote statistical averaging (we discuss a more complicated and realistic scenario that leads to the same general conclusions in Appendix \ref{sec:error}).
 The mean squared error with which $\tilde\rho^V_k$ approximates $\rho^V_k$ is then
@@
 
-The complexity of computing the error scale $\S_V$ is determined by the need to
-\begin{enumerate*}
-\item construct the measurement matrix $M_V$, and
-\item compute the singular values of $M_V$.
-\end{enumerate*}
-These tasks can be drastically simplified by use of the transition operators.
-Specifically, the orthogonality of transition operators with different degrees, i.e.~$\obk{T_{\v v\ell,0}|T_{\v v',\ell',0}} = 0$ for $\ell\ne\ell'$, implies that $M_V$ has a block-diagonal structure in the transition operator basis, with rectangular blocks
+The complexity of computing the error scale $\S_V$ is determined by the need to construct the measurement matrix $M_V$, and compute its singular values.
+These computations can be drastically reduced by use of the transition operators.
+Specifically, the orthogonality of transition operators with different degrees, i.e.~$\obk{T_{\v v\ell,0}|T_{\v v',\ell',0}} = 0$ for $\ell\ne\ell'$, implies that $M_V$ has a block-diagonal structure in the transition operator basis, with rectangular blocks of size $\p{2\ell+1}\times\abs{V}$:
 \begin{align}
   M_{V\ell} \equiv U_\ell \sum_{\v v} \oket{T_{\v v\ell,0}} \bra{\v v},
-  &&
+\end{align}
+where $U_\ell$ is a $\p{2\ell+1}\times d^2$ isometry defined by
+\begin{align}
   U_\ell \equiv \sum_{m=-\ell}^\ell \ket{m} \obra{T_{\ell m}},
 \end{align}
-that we can use to expand
+with $\obra{T_{\ell m}}$ the conjugate transpose of $\oket{T_{\ell m}}$.
+Using the blocks $M_{V\ell}$, we can expand
 \begin{align}
@@
 where $M_{V\ell}^{-1}$ is the pseudoinverse of $M_{V\ell}$.
-By the expansion $T_{\v v\ell,0} = R\p{\v v} T_{\ell,0} R\p{\v v}^\dag$, the $\p{2\ell+1}\times\abs{V}$ matrix $M_{V\ell}$ can be constructed in $O(\abs{V}d^3)$ time, and computing its singular values takes at most $O(\abs{V}d^2)$ time \footnote{The calculation of $\S_{V\ell}$ can be sped up by the fact that its value is independent of $d$ (as long as $d\ge\ell+1$), so we can essentially set $d=\ell+1$ for the purposes of computing $\S_{V\ell}$}.
-Altogether, if we assume that $\abs{V}\sim d$ then computing $\S_V$ to certify the bound in Eq.~\eqref{eq:bound} takes $O(d^5)$ serial or $O(d^4)$ parallel runtime (see Figure [\red{REF}]).
+Altogether, the expansion $T_{\v v\ell,0} = R\p{\v v} T_{\ell,0} R\p{\v v}^\dag$ enables the construction of $M_{V\ell}$ in $O(\abs{V}d^3)$ time, and computing the singular values of $M_{V\ell}$ takes at most $O(\abs{V}d^2)$ time \footnote{The calculation of $\S_{V\ell}$ can be sped up by the fact that its value is independent of $d$ (as long as $d\ge\ell+1$), so we can essentially set $d=\ell+1$ for the purposes of computing $\S_{V\ell}$}.
+If we assume that $\abs{V}\sim d$, then computing $\S_V$ to certify the bound in Eq.~\eqref{eq:bound} takes $O(d^5)$ serial or $O(d^4)$ parallel runtime (see Figure [\red{REF}]).
 [\red{todo: add figure showing runtimes as a function of $d$ with $\abs{V}=2d-1$; modify text appropriately after making the figure}].
 
-[\red{todo: add a careful derivation of the error bound that assumes $n$ measurements of spin projection along each axis into an Appendix.  Briefly discuss the result of this derivation here.}]
+[\red{todo: add a derivation of the error bound that assumes $n$ measurements of spin projection along each axis to Appendix \ref{sec:error}.
+  Briefly discuss these results in this section.}]
 
@@
 The reconstruction error bound in Eq.~\eqref{eq:bound} provides a straightforward answer: the number of measurement axes should be chosen to minimize the product $\epsilon^2\S_V^2$, which bounds the expected error in the estimate $\tilde\rho_V$ of an unknown state $\rho$.
-If there are a total of $N$ spin projection measurements onto $r$ quantization axes, then $\sim N/r$ measurements are used to estimate spin projection onto any given axis.
-The variance in estimates of spin projection is then $\epsilon^2\sim r/N$, or simply $\epsilon^2\sim r$ at a fixed total number of measurements $N$.
+If there are a total of $N$ spin projection measurements onto $r$ quantization axes, then $n=N/r$ measurements are used to estimate spin projection onto any given axis.
+The variance in estimates of spin projection is then $\epsilon^2\sim1/n\sim r$ at a fixed total number of measurements $N$.
 Altogether, the number of measurement axes should be chosen in such a way as to minimize $r\S_V^2$ with $\abs{V}=r$.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{A closer look at measurement error}
+\label{sec:error}
+
+[\red{todo: derive an error bound that relies on the number of spin projection measurements}]
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Transition operator product expansion}

### 2020-11-04 sun_phases/spin_qudit_tomo.tex
major changes made upon realization that the transition operators of degree L transform identically to spin-L particles this realization greatly simplifies much of the text, and reduces the computational complexity of certifying the error bound

--- 2020-11-02 sun_phases/spin_qudit_tomo.tex
@@
 
-\newcommand{\D}{\mathfrak{D}}
-\renewcommand{\dd}{\mathfrak{d}}
 \renewcommand{\d}{\text{d}}
@@
   We conjecture that $r_d^{\t{min}}=2d-1$, which we verify for all $d\le200$ using numerical experiments.
-  We then provide an algorithm with $O\p{d^4}$ runtime for certifying an upper bound on the error with which measurements of spin projection can reconstruct an unknown quantum state.
+  We then provide an algorithm with $O\p{rd^2}$ runtime for certifying an upper bound on the error with which measurements of spin projection along $r$ axes can reconstruct an unknown quantum state.
   Our algorithm motivates a simple randomized tomography protocol, which we use to study the trade-off between the number of measurement axes and state reconstruction error at a fixed total number of measurements [\red{note: this part is not yet complete}].
@@
 For brevity, we will generally treat the value of $d$ (or $s$) as constant but arbitrary throughout this work, and suppress any explicit dependence of quantities or operators such as $T_{\ell m}$ on $d$.
-The transition operators are orthonormal with respect to the trace inner product $\obk{\O|\Q}\equiv\tr\p{\O^\dag\Q}$, and transform nicely under conjugation:
+The transition operators are orthonormal with respect to the trace inner product, and transform nicely under conjugation:
 \begin{align}
@@
 \end{align}
-where $\delta_{kk'}\equiv 1$ if $k=k'$ and $0$ otherwise.
+where for any $d\times d$ matrix $X=\sum_{\mu,\nu} X_{\mu\nu} \op{\mu}{\nu}$ we define the $d^2$-component vector $\oket{X} \equiv \sum_{\mu,\nu} X_{\mu\nu} \ket{\mu\nu}$; $\obra{X}$ is the conjugate transpose of $\oket{X}$; and $\delta_{kk'}\equiv 1$ if $k=k'$ and $0$ otherwise.
 As a consequence, we can expand any density operator $\rho$ in the transition operator basis as
@@
   &&
-  \rho_{\ell m}\equiv\obk{T_{\ell m}|\rho},
+  \rho_{\ell m} \equiv \obk{T_{\ell m}|\rho},
   \label{eq:trans_state}
@@
 
-The transition operators are spherical tensor operators, whose degree $\ell$ is preserved under 3D rotations generated by the spin operators $S_\x,S_\y,S_\z$ (see Appendix \ref{sec:rotations}).
-For any triplet of Euler angles $\v\omega=\p{\alpha_{\v\omega},\beta_{\v\omega},\gamma_{\v\omega}}$, we can therefore define the {\it fundamental} rotation operator
+The transition operators are spherical tensor operators, whose degree ($\ell$) is preserved under 3D rotations generated by the spin operators $S_\x,S_\y,S_\z$.
+Moreover, the degree-$\ell$ transition operators $T_{\ell m}$ transform identically to spin-$\ell$ particles under 3D rotations (see Appendix \ref{sec:rotations}).
+For any triplet of Euler angles $\v\omega=\p{\alpha_{\v\omega},\beta_{\v\omega},\gamma_{\v\omega}}$, we can therefore define the rotation operator
 \begin{align}
   R\p{\v\omega} \equiv e^{-\i\alpha_{\v\omega}S_\z} e^{-\i\beta_{\v\omega}S_\y} e^{-\i\gamma_{\v\omega}S_\z},
-  \label{eq:rot_op}
 \end{align}
-as well as the corresponding {\it adjoint} rotation operator $\R\p{\v\omega}$ defined by $\R\p{\v\omega}\O = R\p{\v\omega} \O R\p{\v\omega}^\dag$, and expand
+and expand
 \begin{align}
   T_{\v\omega\ell m} \equiv
-  \R\p{\v\omega} T_{\ell m}
-  = \sum_{n=-\ell}^\ell \D_{nm}^\ell\p{\v\omega} T_{\ell n}.
+  R\p{\v\omega} T_{\ell m} R\p{\v\omega}^\dag
+  = \sum_{n=-\ell}^\ell D_{nm}^\ell\p{\v\omega} T_{\ell n},
   \label{eq:trans_rot}
 \end{align}
-The adjoint rotation matrix elements $\D_{mn}^\ell\p{\v\omega}\equiv\obk{T_{\ell m}|\R\p{\v\omega}|T_{\ell n}}$ are an analog of the fundamental (Wigner) rotation matrix elements $D_{mn}^\ell\p{\v\omega}\equiv\bk{\ell m|R\p{\v\omega}|\ell n}$ that govern rotations of spin-$\ell$ particles and spherical harmonics $Y_{\ell m}$, where $\ket{\ell m}$ denotes the state of a spin-$\ell$ qudit with spin projection $m$ onto a quantization axis.
-
-It will be useful to think of rotation matrix elements $D^\ell_{mn}$ and $\D^\ell_{mn}$ as scalar functions of Euler angles $\v\omega\in\Omega_3$.
-As functions of $\v\omega$, the fundamental rotation matrix elements $D_{mn}^\ell$ satisfy the orthogonality relations \cite{brown2003rotational}
+where
 \begin{align}
-  \int_{\Omega_3} \d\Omega_3\p{\v\omega}
-  D^{\ell}_{mn}\p{\v\omega}^* D^{\ell'}_{m'n'}\p{\v\omega}
-  = \f{8\pi^2}{2\ell+1} \delta_{\ell\ell'} \delta_{mm'} \delta_{nn'},
-  \label{eq:ortho_full}
+  D^\ell_{mn}\p{\v\omega}
+  \equiv \bk{\ell m|R\p{\v\omega}|\ell n}
 \end{align}
-where integration is performed over all $\v\omega\in\Omega_3\simeq[0,2\pi]\times[0,\pi]\times[0,2\pi]$ with integration measure $\d\Omega_3\p{\v\omega} \equiv \d\alpha_{\v\omega} \d\beta_{\v\omega} \d\gamma_{\v\omega} \sin\beta_{\v\omega}$.
-In particular, the restriction of $D^\ell_{m,0}$ to the sphere $\Omega_2$ at $\gamma_{\v\omega}=0$ recovers (up to complex conjugation) the spherical harmonics $Y_{\ell m}$, with orthogonality relations
-\begin{align}
-  \int_{\Omega_2} \d\Omega_2\p{\v v}
-  D^{\ell}_{m,0}\p{\v v}^* D^{\ell'}_{m',0}\p{\v v}
-  = \f{4\pi}{2\ell+1} \delta_{\ell\ell'} \delta_{mm'},
-  \label{eq:ortho_sphere}
-\end{align}
-where integration is performed over all $\v v\in\Omega_2\simeq[0,2\pi]\times[0,\pi]$ with integration measure $\d\Omega_2\p{\v v} \equiv \d\alpha_{\v v} \d\beta_{\v v} \sin\beta_{\v v}$, and for ease of notation we implicitly associate doublets $\v v=\p{\alpha_{\v v},\beta_{\v v}}$ with the triplet $\p{\alpha_{\v v},\beta_{\v v},0}$, so that for example $R\p{\v v} = R\p{\alpha_{\v v},\beta_{\v v},0}$ and $T_{\v v\ell m} = T_{\p{\alpha_{\v v},\beta_{\v v},0},\ell m}$.
-The same orthogonality relations in Eqs.~\eqref{eq:ortho_full} and \eqref{eq:ortho_sphere} are inherited by the adjoint rotation matrix elements $\D_{mn}^\ell$ through the fact $\R\p{\v\omega}$ is the adjoint action of $R\p{\v\omega}$ (see Appendix \ref{sec:ortho}), and the functions $\set{\D_{m,0}^\ell}$ restricted to $\Omega_2$ are similarly another orthogonal basis for the space $L^2\p{\Omega_2}$ of square-integrable functions on the sphere.
+are (Wigner) rotation matrix elements.
+Throughout this work, we will primarily consider rotations determined by points on the sphere.
+For ease of notation, we will therefore implicitly associate Euler angle doublets $\v v=\p{\alpha_{\v v},\beta_{\v v}}$ with the triplet $\p{\alpha_{\v v},\beta_{\v v},0}$, so that for example $R\p{\v v} = R\p{\alpha_{\v v},\beta_{\v v},0}$ and $T_{\v v\ell m} = T_{\p{\alpha_{\v v},\beta_{\v v},0},\ell m}$.
 
@@
 Our goal is to reconstruct an arbitrary state of a spin qudit from measurements of spin projection onto different quantization axes.
-We are thus nominally restricted to measuring projectors $\Pi_{\v v\mu} \equiv \op{\mu_{\v v}}$, where $\ket{\mu_{\v v}}\equiv R\p{\v v}\ket{\mu}$ is a state with spin projection $\mu$ onto a measurement axis determined by a point $\v v$ on the sphere $\Omega_2$.
+We are thus nominally restricted to measuring projectors $\Pi_{\v v\mu} \equiv \op{\mu_{\v v}}$, where $\ket{\mu_{\v v}}\equiv R\p{\v v}\ket{\mu}$ is a state with spin projection $\mu$ onto a measurement axis determined by a point $\v v$ on the sphere.
 For any fixed axis $\v v$, the sets $\set{\Pi_{\v v\mu}}$ and $\set{T_{\v v\ell,0}}$ are both orthonormal bases for the space of operators that are diagonal in the basis $\set{\ket{\mu_{\v v}}}$.
@@
   T_{\ell m}
-  = \sum_{\v v,k,n} C_{\ell mk}\p{\v v} \D^k_{n,0}\p{\v v} T_{kn}.
+  = \sum_{\v v,k,n} C_{\ell mk}\p{\v v} D^k_{n,0}\p{\v v} T_{kn}.
 \end{align}
@@
 \begin{align}
-  \sum_{\v v} C_{\ell m}\p{\v v} \D^\ell_{n,0}\p{\v v} = \delta_{mn}
+  \sum_{\v v} C_{\ell m}\p{\v v} D^\ell_{n,0}\p{\v v} = \delta_{mn}
   \label{eq:tomo_recovery}
@@
 In fact, the problem of finding suitable axes $V$ and coefficients $C_{\ell m}\p{\v v}$ to satisfy Eq.~\eqref{eq:tomo_recovery}, solving which would enable the recovery of arbitrary spin qudit states, can be mapped onto the well-studied problem of signal recovery on the sphere \cite{mcewen2011novel, mcewen2011sampling, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
-The signal recovery problem can be stated as follows: given a square-integrable function $f\in L^2\p{\Omega_2}$ on the sphere, with the harmonic expansion
+The signal recovery problem can be stated as follows: given a square-integrable function $f$ on the sphere, with the harmonic expansion
 \begin{align}
-  f\p{\v v} = \sum_{\ell,m} f_{\ell m} B_{\ell m}\p{\v v},
+  f\p{\v v} = \sum_{\ell,m} f_{\ell m} Y_{\ell m}\p{\v v},
 \end{align}
-where $f_{\ell m}$ are expansion coefficients and $\set{B_{\ell m}}$ is an orthogonal basis for $L^2\p{\Omega_2}$, find a set of points $V=\set{\v v}\subset\Omega_2$ and associated coefficients $C_{\ell m}\p{\v v}$ with which we can reconstruct $f$, or equivalently its coefficients $f_{\ell m}$, from knowledge of the function's value $f\p{\v v}$ at all points $\v v\in V$; that is
+where $f_{\ell m}$ are expansion coefficients and $Y_{\ell m}$ is the spherical harmonic of degree $\ell$ and order $m$, find a set of points $V=\set{\v v}$ and associated coefficients $C_{\ell m}\p{\v v}$ with which we can reconstruct $f$, or equivalently its coefficients $f_{\ell m}$, from knowledge of the function's value $f\p{\v v}$ at all points $\v v\in V$; that is
 \begin{align}
   f_{\ell m} = \sum_{\v v} C_{\ell m}\p{\v v} f\p{\v v}
-  = \sum_{\v v,k,n} C_{\ell m}\p{\v v} f_{kn} B_{kn}\p{\v v}.
+  = \sum_{\v v,k,n} C_{\ell m}\p{\v v} f_{kn} Y_{kn}\p{\v v}.
 \end{align}
@@
 \begin{align}
-  \sum_{\v v} C_{\ell m}\p{\v v} B_{kn}\p{\v v}
+  \sum_{\v v} C_{\ell m}\p{\v v} Y_{kn}\p{\v v}
   = \delta_{k\ell} \delta_{mn},
@@
 We will refer to Eq.~\eqref{eq:full_recovery} as the {\it full recovery problem}, and Eq.~\eqref{eq:tomo_recovery} as the the {\it reduced recovery problem}.
-Any solution to the full recovery problem with basis functions $B_{\ell m}=\D^\ell_{m,0}$ automatically solves the reduced recovery problem.
+Any solution to the full recovery problem with basis functions $Y_{\ell m}=D^\ell_{m,0}$ automatically solves the reduced recovery problem.
 
-If
-\begin{enumerate*}
-\item the functions $B_{\ell m}$ are indexed by an integer degree $\ell\ge0$ that is preserved under rotations of the sphere and an integer order $m\in\set{-\ell,-\ell-1,\cdots,\ell}$, as with the spherical harmonics $Y_{\ell m}$ and $\D^\ell_{m,0}$, and
-\item the function $f$ is {\it band-limited} at degree $L$, which is to say that $f_{\ell m}=0$ for all $\ell\ge L$,
-\end{enumerate*}
-then the full recovery problem in Eq.~\eqref{eq:full_recovery} is provably solvable with a suitable choice of $\abs{V}=L^2$ points \cite{freeden2008spherical, freeden2018spherical}.
+If the function $f$ is {\it band-limited} at degree $L$, which is to say that $f_{\ell m}=0$ for all $\ell\ge L$, then the full recovery problem in Eq.~\eqref{eq:full_recovery} is provably solvable with a suitable choice of $\abs{V}=L^2$ points \cite{freeden2008spherical, freeden2018spherical}.
 The existence of these solutions to the full recovery problem in turn implies the existence of $d^2$ measurement axes that suffice to reconstruct arbitrary states of $d$-level spin qudit, whose possible states are essentially ``band-limited'' at degree $d$ \footnote{Spin qudit states can be faithfully represented by quasi-probability distributions on a sphere.
 For a $d$-level spin, these quasi-probability distributions are band-limited at degree $d$.}.
+Moreover, for any fixed degree $\ell$, finding solutions to the reduced recovery problem in Eq.~\eqref{eq:tomo_recovery} is equivalent to the recovery of a degree-$\ell$ function $f = \sum_m f_m Y_{\ell m}$, which is possible with $\abs{V}=2\ell+1$ samples \cite{freeden2008spherical}.
+In the case of spin qudit tomography, the degree $\ell$ takes a maximal value of $\ell_{\t{max}}\equiv d-1$, so state recovery requires at least as many measurement axes as there are transition operators with degree $\ell_{\t{max}}$, namely $2\ell_{\t{max}}+1=2d-1$.
 
-Moreover, for any fixed degree $\ell$, finding solutions to the reduced recovery problem in Eq.~\eqref{eq:tomo_recovery} is equivalent to the recovery of a function in the degree-$\ell$ subspace of $L^2\p{\Omega_2}$, which is possible with $\abs{V}=2\ell+1$ samples (equal to the dimension of the degree-$\ell$ subspace) \cite{freeden2008spherical}.
-In the case of spin qudit tomography, the degree $\ell$ takes a maximal value of $\ell_{\t{max}}\equiv d-1$, so state recovery requires at least as many measurement axes as there are transition operators with degree $\ell_{\t{max}}$, namely $2\ell_{\t{max}}+1=2d-1$.
 Since
@@
 
-In principle, any solution to the full recovery problem with a particular set of basis functions, such the spherical harmonics $Y_{\ell m}$, can be readily translated into a solution with a different set of basis functions, such as the adjoint rotation matrix elements $\D^\ell_{m,0}$ (see Appendix \ref{sec:basis_change}).
-This translation enables importing a host of existing signal recovery algorithms \cite{mcewen2011novel, mcewen2011sampling, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality} for the task of spin qudit tomography.
-In practice, spin qudits typically have only a modest dimension $d$, allowing for simpler and optimized tomography protocols that are practical despite worse scaling with $d$.
+In principle, our mapping of the spin qudit tomography problem onto that of signal recovery on the sphere allows us to import a host of existing signal recovery algorithms \cite{mcewen2011novel, mcewen2011sampling, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality} for the task of spin qudit tomography.
+In practice, spin qudits typically have only a modest dimension $d$, allowing for simpler and optimized tomography protocols that are practical despite worse scaling with $d$ (see Section \ref{sec:protocol}).
 
@@
 
-A choice of measurement axes $V=\set{\v v}$ determines a set of projectors $\v\Pi_V \equiv \set{\Pi_{\v v\mu}}$ that will be measured in an experiment.
-By flattening qudit operators $X = \sum_{\mu,\nu} X_{\mu\nu} \op{\mu}{\nu}$ into the $d^2$-component vectors $\oket{X} \equiv \sum_{\mu,\nu} X_{\mu\nu} \ket{\mu\nu}$, we can construct the {\it measurement matrix}
+A set of measurement axes $V=\set{\v v}$ nominally induces a set of projectors $\set{\Pi_{\v v\mu}}$ that will be measured in an experiment.
+By a simple change of basis, measuring these projectors is equivalent to measuring the set of transition operators $\v T_V\equiv\set{T_{\v v\ell,0}}$.
+By flattening the operators in $\v T_V$ into column vectors $\oket{T_{\v\ell,0}}$, we can construct the {\it measurement matrix}
 \begin{align}
-  M_V \equiv \sum_{\v v,\ell} \oket{\Pi_{\v v\mu}} \bra{\v v\mu},
+  M_V \equiv \sum_{\v v,\ell} \oket{T_{\v v\ell,0}} \bra{\v v\ell},
   \label{eq:meas_mat}
 \end{align}
-whose columns are flattenings of the operators in $\v\Pi_V$.
-A necessary and sufficient condition for $V$ to allow for full state tomography is that the projectors in $\v\Pi_V$, or equivalently the columns of $M_V$, to span the entire ($d^2$-dimensional) space of operators on a $d$-level spin qudit.
+A necessary and sufficient condition for $V$ to allow for full state tomography is that the transition operators in $\v T_V$, or equivalently the columns of $M_V$, span the entire ($d^2$-dimensional) space of operators on a $d$-level spin qudit.
 In this case $M_V$ must be full rank, with $d^2$ nonzero singular values.
@@
 \begin{align}
-  Q^V_k \equiv \sum_j q^V_{kj} \Pi_j,
+  Q^V_k \equiv \sum_j q^V_{kj} T_j,
   &&
@@
 \end{align}
-where for shorthand we use a combined index $j=\p{\v v_j,\mu_j}$ to specify both a measurement axis $\v v_j$ and a spin projection $\mu_j$, which identify the projector operator $\Pi_j\equiv \Pi_{\v v_j\mu_j}$.
+where for shorthand we use a combined index $j=\p{\v v_j,\mu_j}$ to specify both a measurement axis $\v v_j$ and a spin projection $\mu_j$, which identify the transition operator $T_j\equiv T_{\v v_j\mu_j}$.
 These operators allow us to expand any state $\rho$ of a $d$-level spin qudit in the form
@@
 \end{align}
-Given empirical estimates $E_j$ of the expectation values $\bk{\Pi_j}_\rho$, an empirical estimate $\tilde\rho_V$ of $\rho$ is then
+Given empirical estimates $E_j$ of the expectation values $\bk{T_j}_\rho$, an empirical estimate $\tilde\rho_V$ of $\rho$ is then
 \begin{align}
@@
   \tilde\rho^V_k \equiv \sum_j q^V_{kj} E_j
-  \approx \sum_j q^V_{kj} \bk{\Pi_j}_\rho
+  \approx \sum_j q^V_{kj} \bk{T_j}_\rho
   = \bk{Q^V_k}_\rho
@@
 The singular values of $M_V$ allow us to make concrete statements about the statistical error between the empirical estimate $\tilde\rho_V$ and the true state $\rho$.
-Assume, for example, that the estimates $E_j$ are equal to $\bk{\Pi_j}_\rho$ up to uncorrelated noise with variance at most $\epsilon^2$, i.e.
+Assume, for example, that the estimates $E_j$ are equal to $\bk{T_j}_\rho$ up to uncorrelated noise with variance at most $\epsilon^2$:
 \begin{align}
-  E_j = \bk{\Pi_j}_\rho + \epsilon_j,
+  E_j = \bk{T_j}_\rho + \epsilon_j,
   &&
@@
 
-The complexity of computing the error scale $\S_V$ is determined by the need to
-\begin{enumerate*}
-\item construct the measurement matrix $M_V$, and
-\item compute the singular values of $M_V$.
-\end{enumerate*}
-These tasks can be drastically simplified by use of the transition operators.
-Specifically, the orthogonality of transition operators with different degrees, i.e.~$\obk{T_{\v v\ell,0}|T_{\v v',\ell',0}} = 0$ for $\ell\ne\ell'$, implies that $M_V$ has a block-diagonal structure in the transition operator basis, with rectangular blocks
+The complexity of computing the error scale $\S_V$ is determined by the need to construct and compute the singular values of the measurement matrix $M_V$.
+This task can be greatly simplified by the fact that the degree $\ell$ of the transition operator $T_{\ell,0}$ is preserved under rotations, which implies that the unitary
 \begin{align}
-  M_{V\ell} \equiv U_\ell \sum_{\v v} \oket{T_{\v v\ell,0}} \bra{\v v},
-  &&
-  U_\ell \equiv \sum_{m=-\ell}^\ell \ket{m} \obra{T_{\ell m}},
+  U \equiv \sum_{\ell,m} \ket{\ell m} \obra{T_{\ell m}},
 \end{align}
-that we can use to expand
+block-diagonalizes the measurement matrix:
+\begin{align}
+  U M_V = \sum_\ell \op{\ell} \otimes M_{V\ell},
+\end{align}
+where the $\p{2\ell+1}\times\abs{V}$ blocks are
+\begin{align}
+  M_{V\ell} \equiv \sum_{\v v,m} \ket{m} \obk{T_{\ell m}|T_{\v v\ell,0}} \bra{\v v}
+  = \sum_{\v v,m} D^\ell_{m,0}\p{\v v} \op{m}{\v v}.
+\end{align}
+As the singular values of $M_V$ are invariant under unitary transformations, it follows that
 \begin{align}
@@
 where $M_{V\ell}^{-1}$ is the pseudoinverse of $M_{V\ell}$.
-By the expansion $T_{\v v\ell,0} = R\p{\v v} T_{\ell,0} R\p{\v v}^\dag$, the $\p{2\ell+1}\times\abs{V}$ matrix $M_{V\ell}$ can be constructed in $O(\abs{V}d^3)$ time, and computing its singular values takes at most $O(\abs{V}d^2)$ time \footnote{The calculation of $\S_{V\ell}$ can be sped up by the fact that its value is independent of $d$ (as long as $d\ge\ell+1$), so we can essentially set $d=\ell+1$ for the purposes of computing $\S_{V\ell}$}.
-Altogether, if we assume that $\abs{V}\sim d$ then computing $\S_V$ to certify the bound in Eq.~\eqref{eq:bound} takes $O(d^5)$ serial or $O(d^4)$ parallel runtime (see Figure [\red{REF}]).
+Altogether, constructing the block $M_{V\ell}$ and computing its singular values takes at most $O(\abs{V}d^2)$ time.
+If we assume that $\abs{V}\sim d$, then computing $\S_V$ to certify the bound in Eq.~\eqref{eq:bound} takes $O(d^4)$ serial or $O(d^3)$ parallel runtime (see Figure [\red{REF}]).
 [\red{todo: add figure showing runtimes as a function of $d$ with $\abs{V}=2d-1$; modify text appropriately after making the figure}].
 
-[\red{todo: add a careful derivation of the error bound that assumes $n$ measurements of spin projection along each axis into an Appendix.  Briefly discuss the result of this derivation here.}]
+[\red{todo: add a derivation of the error bound that assumes $n$ measurements of spin projection along each axis to Appendix \ref{sec:error}.
+  Briefly discuss these results in this section.}]
 
@@
 
-Defining
+Denoting the state of a spin-$s$ particle spin spin projection $\mu$ onto a quantization axis by $\ket{s\mu}$, we define
 \begin{align}
-  S_\z \equiv \sum_{\mu=-s}^s \mu \op{\mu},
+  S_\z \equiv \sum_{\mu=-s}^s \mu \op{s\mu},
   &&
   S_\pm \equiv \sum_{\mu=-s}^s
-  \sqrt{s\p{s+1}-\mu\p{\mu\pm1}} \op{\mu\pm1}{\mu},
+  \sqrt{s\p{s+1}-\mu\p{\mu\pm1}} \op{s,\mu\pm1}{s\mu},
 \end{align}
@@
   &&
-  \v S \equiv \p{S_\x,S_\y,S_\z},
+  \v S \equiv \p{S_\x,S_\y,S_\z}.
 \end{align}
-the spin vector $\v S$ generates rotations of a spin-$s$ system in 3D space.
+The spin vector $\v S$ generates rotations of a spin-$s$ system in 3D space.
 Specifically, the operator $e^{-\i\theta\v S\cdot\uv n}$ rotates a spin-$s$ system by an angle $\theta$ about the unit vector $\uv n$.
-Furthermore, observing that $S_\z=T_{1,0}$ and $S_\pm\propto T_{1,\pm1}$, we can use the product operator expansion of the transition operators (see Appendix \ref{sec:trans_prod}), the properties of Clebsch-Gordan coefficients, and a computer algebra system to simplify the commutators
+
+Observing that $S_\z=T_{1,0}$ and $S_\pm\propto T_{1,\pm1}$, we can use the product operator expansion of the transition operators (see Appendix \ref{sec:trans_prod}), the properties of Clebsch-Gordan coefficients, and a computer algebra system to simplify the commutators
 \begin{align}
@@
   &&
-  \sp{S_\pm,T_{\ell m}} = \sqrt{\p{\ell\mp m}\p{\ell\pm m+1}}\, T_{\ell,m\pm1},
+  \sp{S_\pm,T_{\ell m}} = \sqrt{\ell\p{\ell+1}-m\p{m\pm 1}}\, T_{\ell,m\pm1},
 \end{align}
-which implies that $T_{\ell m}$ is a spherical tensor operator, and in particular that its degree $\ell$ is preserved under rotations generated by $\v S$.
+which implies that $T_{\ell m}$ is a spherical tensor operator, whose degree degree $\ell$ is preserved under rotations generated by $\v S$.
+Moreover, the transition operators $T_{\ell m}$ transform identically to spin-$\ell$ particles under the adjoint action of the spin operators $S_\z$ and $S_\pm$.
 For any triplet of Euler angles $\v\omega=\p{\alpha_{\v\omega},\beta_{\v\omega},\gamma_{\v\omega}}$, we can therefore define the fundamental rotation operator
@@
 \end{align}
-as well as the corresponding adjoint rotation operator $\R\p{\v\omega}$ defined by $\R\p{\v\omega}\O = R\p{\v\omega} \O R\p{\v\omega}^\dag$, and expand
+and expand
 \begin{align}
-  \R\p{\v\omega} T_{\v\omega\ell m}
-  = \sum_{n=-\ell}^\ell \D_{nm}^\ell\p{\v\omega} T_{\ell n}.
+  T_{\v\omega\ell m}
+  \equiv R\p{\v\omega} T_{\ell m} R\p{\v\omega}^\dag
+  = \sum_{n=-\ell}^\ell D_{nm}^\ell\p{\v\omega} T_{\ell n},
 \end{align}
-The adjoint rotation matrix elements $\D_{mn}^\ell\p{\v\omega}\in\CC$ are
+where
 \begin{align}
-  \D_{mn}^\ell\p{\v\omega}
-  &\equiv \obk{T_{\ell m}|\R\p{\v\omega}|T_{\ell n}} \\
-  &= \tr\p{T_{\ell m}^\dag e^{-\i\alpha_{\v\omega}S_\z} e^{-\i\beta_{\v\omega}S_\y} e^{-\i\gamma_{\v\omega}S_\z} T_{\ell n}
-    e^{\i\gamma_{\v\omega}S_\z} e^{\i\beta_{\v\omega}S_\y} e^{\i\alpha_{\v\omega}S_\z}} \\
-  &= e^{-\i\alpha_{\v\omega} m} \,\dd_{mn}^\ell\p{\beta_{\v\omega}}\, e^{-\i\gamma_{\v\omega} n},
+  D_{mn}^\ell\p{\v\omega} \equiv \bk{\ell m|R\p{\v\omega}|\ell n}
 \end{align}
-and the {\it reduced} adjoint rotation matrix elements $\dd_{mn}^\ell\p{\beta}\in\RR$ are
-\begin{align}
-  \dd_{mn}^\ell\p{\beta}
-  &\equiv \D_{mn}^\ell\p{0,\beta,0} \\
-  &= \tr\p{T_{\ell m}^\dag e^{-\i\beta S_\y} T_{\ell n} e^{\i\beta S_\y}} \\
-  &= \f{2\ell+1}{2s+1} \sum_{\mu,\nu=-s}^s
-  \bk{s\mu;\ell m|s,\mu+m} \bk{s\nu;\ell n|s,\nu+n} \times
-  \tr\p{\op{\mu}{\mu+m} e^{-\i\beta S_\y}
-    \op{\nu+n}{\nu} e^{\i\beta S_\y}} \\
-  &= \f{2\ell+1}{2s+1} \sum_{\mu,\nu=-s}^s
-  \bk{s\mu;\ell m|s,\mu+m} \bk{s\nu;\ell n|s,\nu+n} \times
-  d_{\mu\nu}^s\p{\beta} d_{\mu+m,\nu+n}^s\p{\beta}.
-\end{align}
-Here $d_{\mu\nu}^s\p{\beta}\equiv D^s_{\mu\nu}\p{0,\beta,0}\in\RR$ are
-matrix elements of a reduced fundamental (Wigner) rotation matrix.
-Note that the above product of reduced fundamental rotation matrix elements can be expanded as \cite{rose1957elementary}
-\begin{align}
-  d_{\mu\nu}^s\p{\beta} d_{\mu+m,\nu+n}^s\p{\beta}
-  = \sum_{L=0}^{2s} \bk{s\mu;s,\mu+m|L,2\mu+m}
-  \bk{s\nu;s,\nu+n|L,2\nu+n} d_{2\mu+m,2\nu+n}^L\p{\beta}.
-\end{align}
+are matrix elements of the rotation operator $R\p{\v\omega}$ for spin-$\ell$ particles.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Rotation matrix elements as orthogonal functions}
-\label{sec:ortho}
+\section{A closer look at measurement error}
+\label{sec:error}
 
-As functions of Euler angle triplets $\v\omega=\p{\alpha_{\v\omega},\beta_{\v\omega},\gamma_{\v\omega}}$, the fundamental (Wigner) rotation matrix elements $D_{mn}^\ell$ and satisfy the orthogonality relations \cite{brown2003rotational}
-\begin{align}
-  \int_{\Omega_3} \d\Omega_3\p{\v\omega}
-  D^{\ell}_{mn}\p{\v\omega}^* D^{\ell'}_{m'n'}\p{\v\omega}
-  = \f{8\pi^2}{2\ell+1} \delta_{\ell\ell'} \delta_{mm'} \delta_{nn'},
-  \label{eq:orthogonal}
-\end{align}
-where integration is performed over all $\v\omega\in\Omega_3\simeq[0,2\pi]\times[0,\pi]\times[0,2\pi]$ with integration measure $\d\Omega_3\p{\v\omega}\equiv\d\alpha_{\v\omega} \d\beta_{\v\omega} \d\gamma_{\v\omega} \sin\beta_{\v\omega}$.
-This orthogonality relation is a consequence of the fact that, for a fixed degree $\ell$, the matrix
-\begin{align}
-  D^\ell \equiv \sum_{m,n} D^\ell_{mn} \op{m}{n}
-\end{align}
-is an irreducible unitary matrix representation of SO(3), where by unitary we mean that
-\begin{align}
-  \sum_k D^\ell_{km}\p{\v\omega}^* D^\ell_{kn}\p{\v\omega} = \delta_{mn}
-  ~\t{for all}~\v\omega\in\Omega_3\simeq\SO(3).
-\end{align}
-Unitarity of $D^\ell$ follows from the fact that the matrix $D^\ell\p{\v\omega}$ is essentially the degree-$\ell$ block of the block-diagonal fundamental rotation matrix $R\p{\v\omega}$, which is itself unitary.
-The Schur orthogonality relations for irreducible unitary representations of compact Lie groups then imply that \cite{cornwell1997group}
-\begin{align}
-  \int_{\Omega_3} \d\Omega_3\p{\v\omega}
-  D^{\ell}_{mn}\p{\v\omega}^* D^{\ell'}_{m'n'}\p{\v\omega}
-  = \f{\abs{\Omega_3}}{\dim\p{D^\ell}}
-  \delta_{\ell\ell'} \delta_{mm'} \delta_{nn'},
-  \label{eq:ortho_full_apndx}
-\end{align}
-where $\dim\p{D^\ell}=2\ell+1$ is the dimension of $D^\ell$, and analogously
-\begin{align}
-  \abs{\Omega_3} \equiv \int_{\Omega_3} \d\Omega_3\p{\v\omega} = 8\pi^2.
-\end{align}
-Orthogonality relations for adjoint matrix elements $\D^\ell_{mn}$ are derived identically from the fact that the matrix
-\begin{align}
-  \D^\ell \equiv \sum_{m,n} \D^\ell_{mn} \op{m}{n}
-\end{align}
-is an irreducible unitary representation of SO(3).
-This property of $\D^\ell$ follows from the fact that it is an adjoint representation of SO(3), which is isomorphic to the standard representation $D^\ell$ \cite{hall2015lie}.
-
-The decomposition $D^\ell_{mn}\p{\v\omega} = e^{-\i\alpha_{\v\omega}m} d^\ell_{mn}\p{\beta_{\v\omega}} e^{-\i\gamma_{\v\omega}n}$, where $d^\ell_{mn}\p{\beta}\equiv D^\ell_{mn}\p{0,\beta,0}$ is
-a matrix element of a reduced fundamental (Wigner) rotation matrix \cite{rose1957elementary}, implies that setting $n'=n$ in in Eq.~\eqref{eq:ortho_full_apndx} makes the integrand independent of $\gamma_{\v\omega}$.
-Evaluating the integral over $\gamma_{\v\omega}$ and dividing through by $2\pi$ then recovers the orthogonality relations
-\begin{align}
-  \int_{\Omega_2} \d\Omega_2\p{\v v}
-  D^{\ell}_{mn}\p{\v v}^* D^{\ell'}_{m'n}\p{\v v}
-  = \f{4\pi}{2\ell+1} \delta_{\ell\ell'} \delta_{mm'},
-  \label{eq:ortho_sphere_apndx}
-\end{align}
-where integration is performed over all $\v v\in\Omega_2\simeq[0,2\pi]\times[0,\pi]$ with integration measure $\d\Omega_2\p{\v v} \equiv \d\alpha_{\v v} \d\beta_{\v v} \sin\beta_{\v v}$.
-A similar decomposition of the adjoint rotation matrix elements $\D^\ell_{mn}\p{\v\omega}$ (see Appendix \ref{sec:rotations}) implies the identical orthogonality relations.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Changing bases for the signal recovery problem}
-\label{sec:basis_change}
-
-Here we show how one solution to the problem of signal recovery on the sphere with a particular basis of functions (such as the spherical harmonics $Y_{\ell m}$) can be translated into a solution with a different basis (such as the adjoint rotation matrix elements $\D^\ell_{m,0}$).
-Consider a square-integrable function on the sphere, $f\in L^2\p{\Omega_2}$, with the harmonic expansion
-\begin{align}
-  f\p{\v v} = \sum_{\ell,m} f_{\ell m} B_{\ell m},
-\end{align}
-where $f_{\ell m}\in\CC$ is a coefficient and $\set{B_{\ell m}}$ is a complete basis for $L^2\p{\Omega_2}$.
-Without loss of generality, we will assume that for a fixed degree $\ell$ the functions $\set{B_{\ell m}}$ span the degree-$\ell$ subspace of $L^2\p{\Omega_2}$, and that these functions satisfy orthogonality relations analogous to those in Eqs.~\eqref{eq:ortho_sphere} and \eqref{eq:ortho_sphere_apndx}.
-A solution to the signal recovery problem amounts to a set of points $V=\set{\v v}\subset\Omega_2$ and coefficients $C_{\ell m}\p{\v v}$ for which
-\begin{align}
-  f_{\ell m} = \sum_{\v v} C_{\ell m}\p{\v v} f\p{\v v}.
-  \label{eq:recovery_solution}
-\end{align}
-Consider now a different basis $\set{\tilde B_{\ell m}}$ for $L^2\p{\Omega_2}$, satisfying the same general assumptions that we made for $\set{B_{\ell m}}$, with which
-\begin{align}
-  f\p{\v v} = \sum_{\ell, m} \tilde f_{\ell m} \tilde B_{\ell m}.
-\end{align}
-Substituting the harmonic expansion of $f\p{\v v}$ in the original basis $\set{B_{\ell m}}$, we find that
-\begin{align}
-  \tilde f_{\ell m} = \int_{\Omega_2} \d\Omega_2\p{\v v}
-  \tilde B_{\ell m}\p{\v v}^* f\p{\v v}
-  = \sum_n G^\ell_{mn} f_{\ell n},
-  &&
-  G^\ell_{mn} \equiv \int_{\Omega_2} \d\Omega_2\p{\v v}
-  \tilde B_{\ell m}\p{\v v}^* B_{\ell n}\p{\v v},
-  \label{eq:basis_change}
-\end{align}
-where integration is performed over all $\v v\in\Omega_2\simeq[0,2\pi]\times[0,\pi]$ with integration measure $\d\Omega_2\p{\v v} \equiv \d\alpha_{\v v} \d\beta_{\v v} \sin\beta_{\v v}$.
-Note that the inner product between functions $\tilde B_{\ell m}\p{\v v}$ and $B_{kn}\p{\v v}$ for $k\ne\ell$ vanishes because these functions live in orthogonal subspaces of $L^2\p{\Omega_2}$ \cite{freeden2008spherical}.
-Substituting Eq.~\eqref{eq:recovery_solution} into Eq.~\eqref{eq:basis_change}, we find that
-\begin{align}
-  \tilde f_{\ell m}
-  = \sum_n G^\ell_{mn} \sum_{\v v} C_{\ell n}\p{\v v} f\p{\v v}
-  = \sum_{\v v} \tilde C_{\ell m}\p{\v v} f\p{\v v},
-  &&
-  \tilde C_{\ell m}\p{\v v} \equiv \sum_n G^\ell_{mn} C_{\ell n}\p{\v v}.
-\end{align}
-The coefficients $\tilde C_{\ell m}\p{\v v}$ thus solve the signal recovery problem with the new basis $\set{\tilde B_{\ell m}}$ for $L^2\p{\Omega_2}$.
+[\red{todo: derive an error bound that relies on the number of spin projection measurements}]
 

### 2020-11-04 sun_phases/spin_qudit_tomo.tex
some cleaning up after reading through the current spin tomo notes

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Spin transition operators and rotation matrices}
+\section{Spin transition operators and 3D rotations}
 \label{sec:transition_ops}
@@
 The transition operators can be physically interpreted as the transition $\ket\psi\to T_{\ell m}\ket\psi$ induced on the state $\ket\psi$ of a spin-$s$ system by the absorption of a spin-$\ell$ particle with spin projection $m$ onto a fixed quantization axis.
-In analogy with the spherical harmonics $Y_{\ell m}$, we will refer to the indices $\ell$ and $m$ respectively as the {\it degree} and {\it order} of $T_{\ell m}$.
+In analogy with the spherical harmonics $Y_{\ell m}$, we will refer to the index $\ell$ as the {\it degree} and $m$ as the {\it order} of $T_{\ell m}$.
 
-The transition operators are spherical tensor operators, whose degree ($\ell$) is preserved under 3D rotations generated by the spin operators $S_\x,S_\y,S_\z$.
+The transition operators are spherical tensor operators, whose degree is preserved under 3D rotations generated by the spin operators $S_\x,S_\y,S_\z$.
 Moreover, the degree-$\ell$ transition operators $T_{\ell m}$ transform identically to spin-$\ell$ particles under 3D rotations (see Appendix \ref{sec:rotations}).
@@
 \end{align}
-and expand
+and expand rotations of the transition operators as
 \begin{align}
@@
 \begin{align}
-  D^\ell_{mn}\p{\v\omega}
-  \equiv \bk{\ell m|R\p{\v\omega}|\ell n}
+  D_{mn}^\ell\p{\v\omega} \equiv \bk{\ell m|R\p{\v\omega}|\ell n}
 \end{align}
@@
 Our goal is to reconstruct an arbitrary state of a spin qudit from measurements of spin projection onto different quantization axes.
-We are thus nominally restricted to measuring projectors $\Pi_{\v v\mu} \equiv \op{\mu_{\v v}}$, where $\ket{\mu_{\v v}}\equiv R\p{\v v}\ket{\mu}$ is a state with spin projection $\mu$ onto a measurement axis determined by a point $\v v$ on the sphere.
+We are thus nominally restricted to measuring projectors $\Pi_{\v v\mu} \equiv \op{\mu_{\v v}}$, where $\ket{\mu_{\v v}}\equiv R\p{\v v}\ket{\mu}$ is a state with spin projection $\mu$ onto a measurement axis $\v v$ determined by a point on the sphere.
 For any fixed axis $\v v$, the sets $\set{\Pi_{\v v\mu}}$ and $\set{T_{\v v\ell,0}}$ are both orthonormal bases for the space of operators that are diagonal in the basis $\set{\ket{\mu_{\v v}}}$.
-By a simple change of operator basis, measuring the projectors $\set{\Pi_{\v v\mu}}$ is therefore equivalent to measuring the transition operators $\set{T_{\v v\ell,0}}$, and provides data on the expectation values $\bk{T_{\v \ell,0}}_\rho$, where $\bk{X}_\rho\equiv\tr\p{\rho X}$.
+By a simple change of operator basis, measuring the projectors $\set{\Pi_{\v v\mu}}$ is therefore equivalent to measuring the transition operators $\set{T_{\v v\ell,0}}$, and provides data on the expectation values $\bk{T_{\v \ell,0}}_\rho$, where $\bk{X}_\rho\equiv\obk{\rho|X}$.
 
-Reconstructing an arbitrary density operator $\rho$ essentially requires us to find a set of axes $V=\set{\v v}$ and associated coefficients $C_{\ell mk}\p{\v v}$ that would allow us to recover any coefficient $\rho_{\ell m}$ of the expansion in Eq.~\eqref{eq:trans_state} through
+Reconstructing an arbitrary density operator $\rho$ essentially requires us to find a set of axes $V=\set{\v v}$ and associated coefficients $C_{\ell mk}\p{\v v}$ that would allow us to recover any coefficient $\rho_{\ell m}$ in expansion in Eq.~\eqref{eq:trans_state} through
 \begin{align}
@@
 \end{align}
-where $f_{\ell m}$ are expansion coefficients and $Y_{\ell m}$ is the spherical harmonic of degree $\ell$ and order $m$, find a set of points $V=\set{\v v}$ and associated coefficients $C_{\ell m}\p{\v v}$ with which we can reconstruct $f$, or equivalently its coefficients $f_{\ell m}$, from knowledge of the function's value $f\p{\v v}$ at all points $\v v\in V$; that is
+where $f_{\ell m}$ are expansion coefficients and $Y_{\ell m}$ is the spherical harmonic of degree $\ell$ and order $m$, find a set of points $V=\set{\v v}$ and associated coefficients $\tilde C_{\ell m}\p{\v v}$ with which we can reconstruct $f$, or equivalently its coefficients $f_{\ell m}$, from knowledge of the function's value $f\p{\v v}$ at all points $\v v\in V$; that is
 \begin{align}
-  f_{\ell m} = \sum_{\v v} C_{\ell m}\p{\v v} f\p{\v v}
-  = \sum_{\v v,k,n} C_{\ell m}\p{\v v} f_{kn} Y_{kn}\p{\v v}.
+  f_{\ell m} = \sum_{\v v} \tilde C_{\ell m}\p{\v v} f\p{\v v}
+  = \sum_{\v v,k,n} \tilde C_{\ell m}\p{\v v} f_{kn} Y_{kn}\p{\v v}.
 \end{align}
@@
 \begin{align}
-  \sum_{\v v} C_{\ell m}\p{\v v} Y_{kn}\p{\v v}
+  \sum_{\v v} \tilde C_{\ell m}\p{\v v} Y_{kn}\p{\v v}
   = \delta_{k\ell} \delta_{mn},
@@
 We will refer to Eq.~\eqref{eq:full_recovery} as the {\it full recovery problem}, and Eq.~\eqref{eq:tomo_recovery} as the the {\it reduced recovery problem}.
-Any solution to the full recovery problem with basis functions $Y_{\ell m}=D^\ell_{m,0}$ automatically solves the reduced recovery problem.
+Due to the fact that $D^k_{n,0}\p{\v v} = \sqrt{\frac{4\pi}{2\ell+1}}\, Y_{kn}\p{\v v}^*$, any solution to the full recovery problem automatically solves the reduced recovery problem by setting $C_{\ell m}\p{\v v} = \sqrt{\frac{2\ell+1}{4\pi}}\, \tilde C_{\ell m}\p{\v v}^*$.
 
@@
 By a simple change of basis, measuring these projectors is equivalent to measuring the set of transition operators $\v T_V\equiv\set{T_{\v v\ell,0}}$.
-By flattening the operators in $\v T_V$ into column vectors $\oket{T_{\v\ell,0}}$, we can construct the {\it measurement matrix}
+By flattening the operators in $\v T_V$ into column vectors $\oket{T_{\v v\ell,0}}$, we can construct the {\it measurement matrix}
 \begin{align}
@@
 \end{align}
-A necessary and sufficient condition for $V$ to allow for full state tomography is that the transition operators in $\v T_V$, or equivalently the columns of $M_V$, span the entire ($d^2$-dimensional) space of operators on a $d$-level spin qudit.
+A necessary and sufficient condition for $V$ to allow for full state tomography is that the measured transition operators in $\v T_V$, or equivalently the columns of $M_V$, span the entire ($d^2$-dimensional) space of operators on a $d$-level spin qudit.
 In this case $M_V$ must be full rank, with $d^2$ nonzero singular values.
@@
 \end{align}
-where for shorthand we use a combined index $j=\p{\v v_j,\mu_j}$ to specify both a measurement axis $\v v_j$ and a spin projection $\mu_j$, which identify the transition operator $T_j\equiv T_{\v v_j\mu_j}$.
+where for shorthand we use a combined index $j=\p{\v v_j,\ell_j}$ to specify both a measurement axis $\v v_j$ and a degree $\ell_j$, which identify the transition operator $T_j\equiv T_{\v v_j\ell_j,0}$.
 These operators allow us to expand any state $\rho$ of a $d$-level spin qudit in the form
@@
 \begin{align}
-  M_{V\ell} \equiv \sum_{\v v,m} \ket{m} \obk{T_{\ell m}|T_{\v v\ell,0}} \bra{\v v}
-  = \sum_{\v v,m} D^\ell_{m,0}\p{\v v} \op{m}{\v v}.
+  M_{V\ell} \equiv \sum_{m,\v v} \ket{m} \obk{T_{\ell m}|T_{\v v\ell,0}} \bra{\v v}
+  = \sum_{m,\v v} D^\ell_{m,0}\p{\v v} \op{m}{\v v}.
 \end{align}
@@
 If $\abs{V}$ is too large for variational optimization to be practical, one can instead generate many random sets of measurement axes, and then choose the set that minimizes $\S_V$.
-After choosing a set of measurement axes $V$, experimentally characterize the expectation values $\bk{\Pi_j}_\rho$.
-If the experimental measurements $E_j$ of $\bk{\Pi_j}_\rho$ are accurate to within an error $\lesssim\epsilon$, then the state $\tilde\rho_V$ defined in Eq.~\eqref{eq:reconstructed_state} is an approximation of the true state $\rho$ to within a 2-norm distance $\lesssim\epsilon\S_V$.
+After choosing a set of measurement axes $V$, experimentally characterize expectation values of spin projections, $\bk{\Pi_{\v v\mu}}_\rho$, which determine the expectation values $\bk{T_{\v v\ell,0}}_\rho$.
+If the experimental estimates $E_{\v v\ell}$ of $\bk{T_{\v v\ell,0}}_\rho$ are accurate to within an error $\lesssim\epsilon$, then the state $\tilde\rho_V$ defined in Eq.~\eqref{eq:reconstructed_state} is an approximation of the true state $\rho$ to within a 2-norm distance $\lesssim\epsilon\S_V$.
 Furthermore, if the empirical estimate $\tilde\rho_V$ has negative eigenvalues, its 2-norm distance from $\rho$ can be further reduced with maximum-likelihood corrections \cite{smolin2012efficient}.
@@
   \sqrt{s\p{s+1}-\mu\p{\mu\pm1}} \op{s,\mu\pm1}{s\mu},
+  \label{eq:spin_ops}
 \end{align}
@@
   \sp{S_\pm,T_{\ell m}} = \sqrt{\ell\p{\ell+1}-m\p{m\pm 1}}\, T_{\ell,m\pm1},
+  \label{eq:spin_trans}
 \end{align}
 which implies that $T_{\ell m}$ is a spherical tensor operator, whose degree degree $\ell$ is preserved under rotations generated by $\v S$.
-Moreover, the transition operators $T_{\ell m}$ transform identically to spin-$\ell$ particles under the adjoint action of the spin operators $S_\z$ and $S_\pm$.
+Moreover, by comparing Eqs.~\eqref{eq:spin_ops} and \eqref{eq:spin_trans} we see that the transition operators $T_{\ell m}$ transform identically to spin-$\ell$ particles under the adjoint action of the spin operators $S_\z$ and $S_\pm$.
 For any triplet of Euler angles $\v\omega=\p{\alpha_{\v\omega},\beta_{\v\omega},\gamma_{\v\omega}}$, we can therefore define the fundamental rotation operator
@@
   D_{mn}^\ell\p{\v\omega} \equiv \bk{\ell m|R\p{\v\omega}|\ell n}
+  = \obk{T_{\ell m}|R\p{\v\omega}\otimes R\p{\v\omega}^*|T_{\ell n}}
 \end{align}

### 2020-11-08 sun_phases/spin_qudit_tomo.tex
some change in notation, rearranging text, and an improved reconstruction error bound

@@
 \renewcommand{\sp}[1]{\left[#1\right]} % square parenthesis
-\renewcommand{\set}[1]{\{#1\}} % curly parenthesis
-\renewcommand{\Set}[1]{\left\{#1\right\}} % curly parenthesis
 \newcommand{\bk}{\Braket} % shorthand for braket notation
@@
 \newcommand{\uv}[1]{\bm{\hat{#1}}} % unit vectors
-\renewcommand{\c}{\cdot} % inner product
 \renewcommand{\i}{\mathrm{i}\mkern1mu} % imaginary unit
+
+% curly brackets for a set
+\renewcommand{\set}[1]{\{#1\}} % curly parenthesis
+\renewcommand{\Set}[1]{\left\{#1\right\}} % curly parenthesis
 
@@
 
-\usepackage{dsfont} % for identity operator
-\newcommand{\1}{\mathds{1}}
-
-\newcommand{\up}{\uparrow}
-\newcommand{\dn}{\downarrow}
 \newcommand{\x}{\text{x}}
@@
 \newcommand{\z}{\text{z}}
+\renewcommand{\d}{\text{d}}
 
-\newcommand{\B}{\mathcal{B}}
 \renewcommand{\O}{\mathcal{O}}
@@
 \renewcommand{\S}{\mathcal{S}}
-\newcommand{\T}{\mathcal{T}}
-\newcommand{\Y}{\mathcal{Y}}
-
-\renewcommand{\d}{\text{d}}
-
-\newcommand{\SO}{\text{SO}}
-
-\newcommand{\CC}{\mathbb{C}}
-\newcommand{\RR}{\mathbb{R}}
-\renewcommand{\SS}{\mathbb{S}}
-\newcommand{\ZZ}{\mathbb{Z}}
 
@@
 \email{mika.perlin@gmail.com}
-% \author{Diego Barberena}
+\author{Diego Barberena}
 \author{Ana Maria Rey}
@@
 We first introduce, in Section \ref{sec:transition_ops}, a basis of operators that have a simple physical interpretation, and transform nicely under rotations in 3D space, thereby respecting the symmetries of the tomography task at hand.
-We then map this quantum tomography problem onto the well-studied classical problem of signal recovery on the sphere in Section \ref{sec:signal_recovery}, thereby importing a host of existing literature and mathematical machinery that has been developed for this task \cite{mcewen2011novel, mcewen2011sampling, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
+We then map this quantum tomography problem onto the well-studied classical problem of signal recovery on the sphere in Section \ref{sec:signal_recovery}, thereby importing a host of existing literature and mathematical machinery that has been developed for this task \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
 In particular, this mapping allows us to bound the minimum number of measurement axes necessary to perform full tomography on a $d$-level spin qudit, $r_d$, by $2d-1\le r_d\le d^2$.
@@
 \label{sec:transition_ops}
+
+[\red{todo: comment on connection between $T_{\ell m}$ and $Y_{\ell m}$}]
 
@@
 
+[\red{todo: add visualization of the signal recovery problem}]
+
 Our goal is to reconstruct an arbitrary state of a spin qudit from measurements of spin projection onto different quantization axes.
@@
 
-In fact, the problem of finding suitable axes $V$ and coefficients $C_{\ell m}\p{\v v}$ to satisfy Eq.~\eqref{eq:tomo_recovery}, solving which would enable the recovery of arbitrary spin qudit states, can be mapped onto the well-studied problem of signal recovery on the sphere \cite{mcewen2011novel, mcewen2011sampling, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
+In fact, the problem of finding suitable axes $V$ and coefficients $C_{\ell m}\p{\v v}$ to satisfy Eq.~\eqref{eq:tomo_recovery}, solving which would enable the recovery of arbitrary spin qudit states, can be mapped onto the well-studied problem of signal recovery on the sphere \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
 The signal recovery problem can be stated as follows: given a square-integrable function $f$ on the sphere, with the harmonic expansion
@@
 Due to the fact that $D^k_{n,0}\p{\v v} = \sqrt{\frac{4\pi}{2\ell+1}}\, Y_{kn}\p{\v v}^*$, any solution to the full recovery problem automatically solves the reduced recovery problem by setting $C_{\ell m}\p{\v v} = \sqrt{\frac{2\ell+1}{4\pi}}\, \tilde C_{\ell m}\p{\v v}^*$.
+In principle, this mapping allows us to import a host of existing signal recovery algorithms \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality} for the task of spin qudit tomography.
+In practice, spin qudits typically have only a modest dimension $d$, which allows for simpler and optimized tomography protocols that are practical despite worse scaling with $d$ (see Section \ref{sec:protocol}).
 
@@
 However, this condition is stronger than that necessary to guarantee full recovery of a $d$-level spin qudit state with $2d-1$ measurement axes.}.
-Unfortunately, we have no general proof of this conjecture, which may require a deeper analysis of fixed-degree signal recovery on the sphere.
-Nonetheless, we have numerically verified that $2d-1$ measurement axes suffice for all $d\le200$ (see Section \ref{sec:axes}).
-
-In principle, our mapping of the spin qudit tomography problem onto that of signal recovery on the sphere allows us to import a host of existing signal recovery algorithms \cite{mcewen2011novel, mcewen2011sampling, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality} for the task of spin qudit tomography.
-In practice, spin qudits typically have only a modest dimension $d$, allowing for simpler and optimized tomography protocols that are practical despite worse scaling with $d$ (see Section \ref{sec:protocol}).
+Unfortunately, we have no general proof of this conjecture, which requires a deeper analysis of fixed-degree signal recovery on the sphere.
+Nonetheless, we have numerically verified that $2d-1$ measurement axes suffice for all $d\le200$ (see Section \ref{sec:bound}).
 
@@
 \end{align}
-Given empirical estimates $E_j$ of the expectation values $\bk{T_j}_\rho$, an empirical estimate $\tilde\rho_V$ of $\rho$ is then
+Given empirical estimates $\tilde T_j$ of the expectation values $\bk{T_j}_\rho$, an empirical estimate $\tilde\rho_V$ of $\rho$ is then
 \begin{align}
@@
 \begin{align}
-  \tilde\rho^V_k \equiv \sum_j q^V_{kj} E_j
+  \tilde\rho^V_k \equiv \sum_j q^V_{kj} \tilde T_j
   \approx \sum_j q^V_{kj} \bk{T_j}_\rho
@@
 The singular values of $M_V$ allow us to make concrete statements about the statistical error between the empirical estimate $\tilde\rho_V$ and the true state $\rho$.
-Assume, for example, that the estimates $E_j$ are equal to $\bk{T_j}_\rho$ up to uncorrelated noise with variance at most $\epsilon^2$:
+Assume, for example, that the estimates $\tilde T_j$ are equal to $\bk{T_j}_\rho$ up to uncorrelated noise with variance at most $\epsilon^2$:
 \begin{align}
-  E_j = \bk{T_j}_\rho + \epsilon_j,
+  \tilde T_j = \bk{T_j}_\rho + \epsilon_j,
   &&
   \bbk{\epsilon_j \epsilon_{j'}} \le \epsilon^2 \delta_{jj'},
+  \label{eq:error_assumption}
 \end{align}
-where $\set{\epsilon_j}$ are independent random variables and we use the double brackets $\bbk{\cdot}$ to denote statistical averaging (we discuss a more complicated and realistic scenario that leads to the same general conclusions in Appendix \ref{sec:error}).
-The mean squared error with which $\tilde\rho^V_k$ approximates $\rho^V_k$ is then
+where $\set{\epsilon_j}$ are independent random variables and we use the double brackets $\bbk{\cdot}$ to denote statistical averaging.
+The assumption in Eq.~\eqref{eq:error_assumption} is reasonable when measurement error is dominated by experimental sources of noise, but breaks down when measurement error is limited by fundamental quantum (shot) noise.
+We relax the assumption of Eq.~\eqref{eq:error_assumption} and perform a proper treatment of measurement error in Appendix \ref{sec:error}.
+Our full treatment of measurement error yields identical conclusions, with an explicit form for $\epsilon$ in terms of the qudit dimension $d$ and the number of spin projection measurements made along every axis $\v v\in V$.
+To simplify our analysis, we therefore maintain the assumption of Eq.~\eqref{eq:error_assumption} for now, in which case the mean squared error with which $\tilde\rho^V_k$ approximates $\rho^V_k$ is
 \begin{align}
@@
 \end{align}
-Using the fact that the operators $Q^V_k$ are orthonormal, we can therefore bound the mean squared 2-norm (Euclidean) distance between $\tilde\rho_V$ and $\rho$ as
+Using the fact that the operators $Q^V_k$ are orthonormal, we can therefore bound the mean squared (Euclidean) distance (i.e.~the distance induced by the trace inner product) between $\tilde\rho_V$ and $\rho$ as
 \begin{align}
-  \Bbk{\norm{\tilde\rho_V-\rho}_2^2}
+  \Bbk{\norm{\tilde\rho_V-\rho}^2}
   = \sum_k \Bbk{\abs{\tilde\rho^V_k - \rho^V_k}^2}
@@
 \begin{align}
-  \S_V \equiv \norm{M_V^{-1}}_2 = \sqrt{\sum_k\f1{\p{M^V_k}^2}},
+  \S_V \equiv \norm{M_V^{-1}} = \sqrt{\sum_k\f1{\p{M^V_k}^2}},
   \label{eq:scale}
 \end{align}
-is the 2-norm of the pseudoinverse $M_V^{-1}$ of $M_V$.
-Note that the bound in Eq.~\eqref{eq:bound} only holds if the measurement matrix $M_V$ is nonsingular, as otherwise some of the operators $Q^V_k$ that we used to expand $\rho$ and $\tilde\rho_V$ are undefined, reflecting the fact that if $M_V$ is singular then measuring spin projections along all axes in $V$ does not provide sufficient information to reconstruct arbitrary quantum states.
+is the (Hilbert-Schmidt, or Frobenius) norm of the right inverse $M_V^{-1}$ of $M_V$.
+Note that the error scale $\S_V$ diverges if $M_V$ is singular, in which case measuring spin projections along all axes in $V$ does not provide sufficient information to reconstruct arbitrary quantum states.
 
@@
   &&
-  \S_{V\ell} \equiv \norm{M_{V\ell}^{-1}}_2,
+  \S_{V\ell} \equiv \norm{M_{V\ell}^{-1}},
 \end{align}
-where $M_{V\ell}^{-1}$ is the pseudoinverse of $M_{V\ell}$.
-Altogether, constructing the block $M_{V\ell}$ and computing its singular values takes at most $O(\abs{V}d^2)$ time.
+where $M_{V\ell}^{-1}$ is the right inverse of $M_{V\ell}$.
+Altogether, constructing the block $M_{V\ell}$ and computing its singular values takes at most $O(\abs{V}d^2)$ time (see Appendix \ref{sec:compute} for additional details on efficiently computing $\S_{V\ell}$ [\red{todo: write tutorial}]).
 If we assume that $\abs{V}\sim d$, then computing $\S_V$ to certify the bound in Eq.~\eqref{eq:bound} takes $O(d^4)$ serial or $O(d^3)$ parallel runtime (see Figure [\red{REF}]).
 [\red{todo: add figure showing runtimes as a function of $d$ with $\abs{V}=2d-1$; modify text appropriately after making the figure}].
+For comparison, we note that existing algorithms for signal recovery on the sphere \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality} similarly have $O(d^3)$ parallel runtime with $\abs{V}\gtrsim d$.
 
@@
   Briefly discuss these results in this section.}]
+
+[\red{todo: add a state-dependent error bound}]
 
@@
 
-The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and variationally optimize the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the error scale $\S_V$ defined in Eq.~\eqref{eq:scale}.
-If $\abs{V}$ is too large for variational optimization to be practical, one can instead generate many random sets of measurement axes, and then choose the set that minimizes $\S_V$.
+The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and optimize (using any standard minimization algorithm) the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the error scale $\S_V$ defined in Eq.~\eqref{eq:scale}.
+If $\abs{V}$ is too large for optimization to be practical, one can instead generate many random sets of measurement axes, and then choose the set with the smallest error scale $\S_V$.
 After choosing a set of measurement axes $V$, experimentally characterize expectation values of spin projections, $\bk{\Pi_{\v v\mu}}_\rho$, which determine the expectation values $\bk{T_{\v v\ell,0}}_\rho$.
-If the experimental estimates $E_{\v v\ell}$ of $\bk{T_{\v v\ell,0}}_\rho$ are accurate to within an error $\lesssim\epsilon$, then the state $\tilde\rho_V$ defined in Eq.~\eqref{eq:reconstructed_state} is an approximation of the true state $\rho$ to within a 2-norm distance $\lesssim\epsilon\S_V$.
-Furthermore, if the empirical estimate $\tilde\rho_V$ has negative eigenvalues, its 2-norm distance from $\rho$ can be further reduced with maximum-likelihood corrections \cite{smolin2012efficient}.
+If the experimental estimates of $\bk{T_{\v v\ell,0}}_\rho$ are accurate to within an error $\lesssim\epsilon$, then the state $\tilde\rho_V$ defined in Eq.~\eqref{eq:reconstructed_state} is an approximation of the true state $\rho$ to within a (Euclidean) distance $\lesssim\epsilon\S_V$.
+Furthermore, if the empirical estimate $\tilde\rho_V$ has negative eigenvalues, its distance from $\rho$ can be further reduced with maximum-likelihood corrections \cite{smolin2012efficient}.
 
@@
 
-Observing that $S_\z=T_{1,0}$ and $S_\pm\propto T_{1,\pm1}$, we can use the product operator expansion of the transition operators (see Appendix \ref{sec:trans_prod}), the properties of Clebsch-Gordan coefficients, and a computer algebra system to simplify the commutators
+Observing that $S_\z=T_{1,0}$ and $S_\pm\propto T_{1,\pm1}$, we can use the product operator expansion of the transition operators (see Appendix \ref{sec:trans_prod}), properties of Clebsch-Gordan coefficients, properties of Wigner $6$-$j$ symbols, and a computer algebra system to simplify the commutators
 \begin{align}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{A closer look at measurement error}
+\section{Relaxing assumptions of the reconstruction error bound}
 \label{sec:error}
 
-[\red{todo: derive an error bound that relies on the number of spin projection measurements}]
+In Section \ref{sec:bound} of the main text, we provided a reconstruction error bound using the assumption of Eq.~\eqref{eq:error_assumption}, namely that expectation values derived from spin projection measurements can be estimated up to uncorrelated errors with maximal variance $\epsilon^2$.
+This assumption is reasonable if measurement error is dominated by experimental sources of noise, and it yields a simple derivation of the reconstruction bound in Eq.~\eqref{eq:bound}.
+Nonetheless, there are two problems with the assumption of Eq.~\eqref{eq:error_assumption}:
+\begin{enumerate*}
+\item there is no a priori guarantee for the value of $\epsilon$, which must be inferred from experimental outcomes, and
+\item the assumption that all errors are uncorrelated is unjustified (and generally false).
+\end{enumerate*}
+Here, we relax the assumption of Eq.~\eqref{eq:error_assumption} and derive an explicit error bound in terms of the qudit dimension $d$ and the number of spin projection measurements made along every measurement axis.
+
+To this end, we fix a particular set of measurement axes $V$, and consider performing $n$ measurements of spin projection along every axis $\v v\in V$, for a total of $N=\abs{V}\times n$ measurements.
+Such a procedure is equivalent to making $N$ local measurements of the $N$-fold product state $\rho^{\otimes N}$.
+For convenience, we index the tensor factors of $\rho^{\otimes N}$ by the integers $\p{i,j}$, with $i\in\set{1,2,\cdots,\abs{V}}$ specifying a measurement axis $\v v_i\in V$ and $j\in\set{1,2,\cdots,n}$ specifying the copy of $\rho$ prepared for the $j$-th measurement spin projection along a particular axis.
+We then define the projectors $\Pi_{i\mu}\equiv\op{\mu_{\v v_i}}$ onto states $\ket{\mu_{\v v_i}}$ with definite spin projection $\mu$ onto axis $\v v_i\in V$, and define $\Pi_{i\mu}^{(j)}$ to be an $N$-qudit operator with $\Pi_{i\mu}$ on the $\p{i,j}$-th tensor factor and the identity elsewhere.
+We denote the experimental outcome of measuring $\Pi_{i\mu}$ in the $\p{i,j}$-th copy of $\rho$ by $\tilde \Pi_{i\mu}^{(j)}\in\set{0,1}$.
+In other words, $\tilde \Pi_{i\mu}^{(j)}$ is the ``single-shot estimate'' of $\Pi_{i\mu}$, with $\tilde \Pi_{i\mu}^{(j)}=1$ if outcome $\mu$ was observed on the $\p{i,j}$-th experimental trial, and $\tilde \Pi_{i\mu}^{(j)}=0$ otherwise.
+An empirical estimate of the expectation value $\bk{\Pi_{i\mu}}_\rho$ is then provided by the fraction of times that outcome $\mu$ was observed when measuring spin projection along axis $\v v_i$, that is
+\begin{align}
+  \tilde \Pi_{i\mu} \equiv \f1n \sum_{j=1}^n \tilde \Pi_{i\mu}^{(j)}
+  \approx \f1n \sum_{j=1}^n \tr\p{\Pi_{i\mu}^{(j)} \rho^{\otimes N}}
+  = \tr\p{\Pi_{i\mu} \rho}.
+\end{align}
+For reasons that will become clear shortly, it will be useful to think of $\tilde \Pi_{i\mu}$ as an empirical estimate of $\bk{\bar \Pi_{i\mu}}_{\rho^{\otimes N}}$, where
+\begin{align}
+  \bar \Pi_{i\mu} \equiv \f1n \sum_{j=1}^n \Pi_{i\mu}^{(j)}
+  \label{eq:mean_proj}
+\end{align}
+is the average of $\Pi_{i\mu}$ applied to all copies of $\rho$ for which spin projection is measured along the axis $\v v_i$.
+
+\subsection{Covariance of errors in the spin-projection basis}
+
+Due to finite sampling error (i.e.~shot noise), an empirical estimate $\tilde\O$ of the observable $\O$ will have statistical error with mean zero:
+\begin{align}
+  \epsilon_\O \equiv \tilde\O - \bk{\O},
+\end{align}
+where the single brackets $\bk{\cdot}$ denote an expectation value with respect to the measured quantum state.
+On average, this statistical error will be zero, which is to say that
+\begin{align}
+  \bbk{\epsilon_\O} = \Bbk{\tilde\O - \bk{\O}}
+  = \bk{\O - \bk{\O}}
+  = 0,
+\end{align}
+where the double brackets $\bbk{\cdot}$ to denote statistical averaging over experimental trials.
+However, the covariance between statistical errors $\epsilon_\O$ and $\epsilon_\Q$ on the empirical estimates $\tilde\O$ and $\tilde\Q$ of observables $\O$ and $\Q$ is
+\begin{align}
+  \bbk{\epsilon_\O \epsilon_\Q}
+  = \Bbk{\p{\tilde\O - \bk{\O}} \p{\tilde\Q - \bk{\Q}}}
+  = \bk{\p{\O - \bk{\O}} \p{\Q - \bk{\Q}}}
+  = \bk{\O\Q} - \bk{\O} \bk{\Q}.
+\end{align}
+In the context of spin qudit tomography, we can therefore define the statistical error
+\begin{align}
+  \epsilon_{i\mu}
+  \equiv \tilde \Pi_{i\mu} - \bk{\Pi_{i\mu}}_{\rho}
+  = \tilde \Pi_{i\mu} - \bk{\bar\Pi_{i\mu}}_{\rho^{\otimes N}}
+\end{align}
+in the empirical estimate of $\bk{\Pi_{i\mu}}_{\rho}$ or $\bk{\bar\Pi_{i\mu}}_{\rho^{\otimes N}}$, and use Eq.~\eqref{eq:mean_proj} to expand
+\begin{align}
+  \bbk{\epsilon_{i\mu} \epsilon_{i'\mu'}}
+  = \bk{\bar\Pi_{i\mu} \bar\Pi_{i'\mu'}}_{\rho^{\otimes N}}
+  - \bk{\bar\Pi_{i\mu}}_{\rho^{\otimes N}}
+  \bk{\bar\Pi_{i'\mu'}}_{\rho^{\otimes N}}
+  = \f1{n^2} \sum_{j,j'=1}^n
+  \sp{\bk{\Pi_{i\mu}^{(j)} \Pi_{i'\mu'}^{(j')}}_{\rho^{\otimes N}}
+    - \bk{\Pi_{i\mu}^{(j)}}_{\rho^{\otimes N}}
+    \bk{\Pi_{i'\mu'}^{(j')}}_{\rho^{\otimes N}}}.
+  \label{eq:proj_cov_start}
+\end{align}
+If $\p{i,j}\ne\p{i',j'}$, then $\Pi_{i\mu}^{(j)}$ and $\Pi_{i'\mu'}^{(j')}$ address different tensor factors of the product state $\rho^{\otimes N}$, so the expectation value of their product factorizes due to the identity $\tr\sp{\p{A\otimes B}\p{A'\otimes B'}} = \tr\p{AA'}\times\tr\p{BB'}$.
+The terms in Eq.~\eqref{eq:proj_cov_start} with $\p{i,j}\ne\p{i',j'}$ therefore vanish, so
+\begin{align}
+  \bbk{\epsilon_{i\mu} \epsilon_{i'\mu'}}
+  &= \delta_{ii'} \times \f1{n^2} \sum_{j=1}^n \sp{\bk{\Pi_{i\mu}^{(j)} \Pi_{i\mu'}^{(j)}}_{\rho^{\otimes N}}
+    - \bk{\Pi_{i\mu}^{(j)}}_{\rho^{\otimes N}}
+    \bk{\Pi_{i\mu'}^{(j)}}_{\rho^{\otimes N}}} \\
+  &= \delta_{ii'} \times \f1n \sp{\bk{\Pi_{i\mu} \Pi_{i\mu'}}_\rho
+    - \bk{\Pi_{i\mu}}_\rho \bk{\Pi_{i\mu'}}_\rho} \\
+  &= \delta_{ii'} \times \f1n \sp{\delta_{\mu\mu'} p_\mu^{(i)} - p_\mu^{(i)} p_{\mu'}^{(i)}},
+\end{align}
+where $p_\mu^{(i)} \equiv \bk{\Pi_{i\mu}}_\rho = \bk{\mu_{\v v_i}|\rho|\mu_{\v v_i}}$ are the diagonal entries of $\rho$ in the basis of spin projections along axis $\v v_i$.
+
+\subsection{Covariance of errors in the transition operator basis}
+
+Rather than the statistical errors $\epsilon_{i\mu} \equiv \tilde\Pi_{i\mu} - \bk{\Pi_{i\mu}}_\rho$ in the estimates $\tilde\Pi_{i\mu}$ of the projectors $\Pi_{i\mu}$, we now consider the statistical errors $\epsilon_{i\ell} \equiv \tilde T_{i\ell} - \bk{T_{i\ell}}_\rho$ in the estimates $\tilde T_{i\ell}$ of the transition operators $T_{i\ell} \equiv T_{\v v_i\ell,0}$.
+We can expand the transition operators $T_{i\ell}$ as a sum over projectors $\Pi_{i\mu}$ as
+\begin{align}
+  T_{i\ell} = \sum_\mu t_{\ell\mu} \Pi_{i\mu},
+  &&
+  t_{\ell\mu} \equiv \bk{\mu|T_{\ell,0}|\mu},
+\end{align}
+and likewise $\tilde T_{i\ell} = \sum_\mu t_{\ell\mu} \tilde\Pi_{i\mu}$.
+The covariances between statistical errors in the transition operator basis are then
+\begin{align}
+  \bbk{\epsilon_{i\ell} \epsilon_{i'\ell'}}
+  = \sum_{\mu,\mu'} t_{\ell\mu} t_{\ell'\mu'}
+  \bbk{\epsilon_{i\mu} \epsilon_{i'\mu'}}
+  = \delta_{ii'} \times \f1n \sum_{\mu,\mu'} t_{\ell\mu} t_{\ell'\mu'}
+  \sp{\delta_{\mu\mu'} p_\mu^{(i)} - p_\mu^{(i)} p_{\mu'}^{(i)}}.
+\end{align}
+In order to establish an upper bound on reconstruction error for an unknown quantum state, we need an upper bound on the variance $\bbk{\epsilon_{i\ell}^2}$ in particular.
+To this end, for any multinomial probability distribution $p\equiv\p{p_1,p_2,\cdots,p_d}$ we define the function
+\begin{align}
+  f_\ell\p{p} \equiv \sum_{\mu,\mu'} t_{\ell\mu} t_{\ell\mu'}
+  \p{\delta_{\mu\mu'} p_\mu - p_\mu p_{\mu'}}
+  = \sum_\mu t_{\ell\mu}^2 p_\mu - \p{\sum_\mu t_{\ell\mu} p_\mu}^2
+  \le \sum_\mu t_{\ell\mu}^2 p_\mu
+  \le \sum_\mu t_{\ell\mu}^2 = 1,
+\end{align}
+which implies that
+\begin{align}
+  \bbk{\epsilon_{i\ell}^2} = \f1n f_\ell\p{p^{(i)}} \le \f1n.
+  \label{eq:variance_bound}
+\end{align}
+
+\subsection{An improved reconstruction error bound}
+
+Equipped with the bound on variances in Eq.~\eqref{eq:variance_bound}, we again consider the block-diagonalized form of the measurement matrix derived in Section \ref{sec:bound}:
+\begin{align}
+  U M_V = \sum_\ell \op{\ell} \otimes M_{V\ell},
+  &&
+  M_{V\ell} = \sum_{m,\v v} D^\ell_{m,0} \op{m}{\v v}.
+\end{align}
+The block diagonal structure of this matrix implies that, rather than an integer index $k\in\set{1,2,\cdots,d^2}$, we can identify the singular values and right singular vectors of $UM_V$ (and equivalently $M_V$) by a block index $\ell\in\set{0,1,\cdots,d-1}$ and a sub-index $m=\set{-\ell,-\ell+1,\cdots,\ell}$.
+Making the replacement $k\to\p{\ell,m}$, we can thus repeat the derivations through Eq.~\eqref{eq:bound} of Section \ref{sec:bound} without the assumption in Eq.~\eqref{eq:error_assumption}, finding that
+\begin{align}
+  \Bbk{\abs{\tilde\rho_{\ell m}^V-\rho_{\ell m}^V}^2}
+  = \sum_{i,i'} q_{\ell m i}^V \p{q_{\ell m i'}^V}^*
+  \bbk{\epsilon_{i\ell} \epsilon_{i'\ell}}
+  = \sum_i \abs{q_{\ell m i}^V}^2
+  \bbk{\epsilon_{i\ell}^2}
+  \le \f1n \sum_i \abs{q_{\ell m i}^V}^2
+  = \f1n \times \f1{\p{M_{\ell m}^V}^2}.
+\end{align}
+Altogether, we thus find that $n$ measurements of spin projection along every axis $\v v\in V$ allows us to construct an empirical estimate $\tilde\rho_V$ of the unknown state $\rho$ that is accurate to within a mean squared distance
+\begin{align}
+  \Bbk{\norm{\tilde\rho_V-\rho}^2} = \sum_{\ell,k}
+  \Bbk{\abs{\tilde\rho_{\ell k}^V-\rho_{\ell k}^V}^2}
+  \le \f{\S_V^2}{n},
+\end{align}
+where $\S_V$ is the same error scale as in Section \ref{sec:bound} of the main text.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Computing the error scale}
+\label{sec:compute}
+
+[\red{todo: write a short tutorial on computing the error scale}]
 

### 2020-11-08 sun_phases/spin_qudit_tomo.tex
tighten upper bound on reconstruction error

@@
 \begin{align}
-  f_\ell\p{p} \equiv \sum_{\mu,\mu'} t_{\ell\mu} t_{\ell\mu'}
+  g_\ell\p{p} \equiv \sum_{\mu,\mu'} t_{\ell\mu} t_{\ell\mu'}
   \p{\delta_{\mu\mu'} p_\mu - p_\mu p_{\mu'}}
-  = \sum_\mu t_{\ell\mu}^2 p_\mu - \p{\sum_\mu t_{\ell\mu} p_\mu}^2
-  \le \sum_\mu t_{\ell\mu}^2 p_\mu
-  \le \sum_\mu t_{\ell\mu}^2 = 1,
+  = \sum_\mu t_{\ell\mu}^2 p_\mu - \p{\sum_\mu t_{\ell\mu} p_\mu}^2.
+\end{align}
+Ideally, we would maximize this function over all valid probability distributions, but doing so is rather difficult.
+Instead, we can find a loose upper bound by ignoring the negative terms:
+\begin{align}
+  g_\ell\p{p} \le \sum_\mu t_{\ell\mu}^2 p_\mu
+  \le \max_\mu t_{\ell\mu}^2
+  = \f{2\ell+1}{d} \max_\mu \bk{s\mu;\ell,0|s\mu}^2
+  \le \f{2\ell+1}{d},
 \end{align}
@@
 \begin{align}
-  \bbk{\epsilon_{i\ell}^2} = \f1n f_\ell\p{p^{(i)}} \le \f1n.
+  \bbk{\epsilon_{i\ell}^2} = \f1n g_\ell\p{p^{(i)}} \le \f{2\ell+1}{nd}.
   \label{eq:variance_bound}
 \end{align}
+Note that $t_{0,\mu}=1/\sqrt{d}$, so $f_0\p{p}=0$ and $\bbk{\epsilon_{i,0}^2}=0$.
 
@@
 The block diagonal structure of this matrix implies that, rather than an integer index $k\in\set{1,2,\cdots,d^2}$, we can identify the singular values and right singular vectors of $UM_V$ (and equivalently $M_V$) by a block index $\ell\in\set{0,1,\cdots,d-1}$ and a sub-index $m=\set{-\ell,-\ell+1,\cdots,\ell}$.
-Making the replacement $k\to\p{\ell,m}$, we can thus repeat the derivations through Eq.~\eqref{eq:bound} of Section \ref{sec:bound} without the assumption in Eq.~\eqref{eq:error_assumption}, finding that
+Making the replacement $k\to\p{\ell,m}$, we can thus repeat the derivations through Eq.~\eqref{eq:bound} of Section \ref{sec:bound} without the assumption of Eq.~\eqref{eq:error_assumption}, finding that
 \begin{align}
@@
   \bbk{\epsilon_{i\ell}^2}
-  \le \f1n \sum_i \abs{q_{\ell m i}^V}^2
-  = \f1n \times \f1{\p{M_{\ell m}^V}^2}.
+  \le \f{2\ell+1}{nd} \p{1-\delta_{\ell,0}}.
 \end{align}
-Altogether, we thus find that $n$ measurements of spin projection along every axis $\v v\in V$ allows us to construct an empirical estimate $\tilde\rho_V$ of the unknown state $\rho$ that is accurate to within a mean squared distance
+Altogether, we thus find that $n$ measurements of spin projection along every axis $\v v\in V$ allows us to construct an empirical estimate $\tilde\rho_V$ of an unknown state $\rho$ that is accurate to within a mean squared distance
 \begin{align}
-  \Bbk{\norm{\tilde\rho_V-\rho}^2} = \sum_{\ell,k}
-  \Bbk{\abs{\tilde\rho_{\ell k}^V-\rho_{\ell k}^V}^2}
-  \le \f{\S_V^2}{n},
+  \Bbk{\norm{\tilde\rho_V-\rho}^2}
+  = \sum_{\ell,m} \Bbk{\abs{\tilde\rho_{\ell m}^V-\rho_{\ell m}^V}^2}
+  \le \f1{nd} \sum_{\ell>0} \p{2\ell+1} \S_{V\ell}^2,
 \end{align}
-where $\S_V$ is the same error scale as in Section \ref{sec:bound} of the main text.
+where
+\begin{align}
+  \S_{V\ell} \equiv \norm{M_{V\ell}^{-1}}
+  = \sqrt{\sum_m \p{M^V_{\ell m}}^{-2}}
+\end{align}
+is the (Hilbert-Schmidt, or Frobenius) norm of the right inverse of $M_{V\ell}$.
 

### 2020-11-08 sun_phases/spin_qudit_tomo.tex
further tightened the error bound

@@
 \end{align}
-where $\bk{s\mu;\ell m|s\nu}$ is a Clebsh-Gordan coefficient that enforces $\ell\in\set{0,1,\cdots,2s}$ and $m\in\set{-\ell,-\ell+1,\cdots,\ell}$, such that there are $d^2$ operators in total.
+where $\bk{s\mu;\ell m|s\nu}$ is a Clebsh-Gordan coefficient that enforces $\ell\in\set{0,1,\cdots,d-1}$ and $m\in\set{-\ell,-\ell+1,\cdots,\ell}$, such that there are $d^2$ operators in total.
 For brevity, we will generally treat the value of $d$ (or $s$) as constant but arbitrary throughout this work, and suppress any explicit dependence of quantities or operators such as $T_{\ell m}$ on $d$.
@@
   \p{\delta_{\mu\mu'} p_\mu - p_\mu p_{\mu'}}
-  = \sum_\mu t_{\ell\mu}^2 p_\mu - \p{\sum_\mu t_{\ell\mu} p_\mu}^2.
+  = \sum_\mu t_{\ell\mu}^2 p_\mu - \p{\sum_\mu t_{\ell\mu} p_\mu}^2
+  = \sigma_p^2\p{t_\ell},
 \end{align}
-Ideally, we would maximize this function over all valid probability distributions, but doing so is rather difficult.
-Instead, we can find a loose upper bound by ignoring the negative terms:
+where $t_\ell \equiv \p{t_{\ell,-s},t_{\ell,-s+1},\cdots,t_{\ell s}}$ is a vector of all $t_{\ell\mu}$, and
+\begin{align}
+  \sigma_p^2\p{X}
+  \equiv \sum_\mu p_\mu \p{X_\mu - \sum_\nu p_\nu X_\nu}^2
+  = \sum_\mu p_\mu X_\mu^2 - \p{\sum_\mu p_\mu X_\mu}^2
+\end{align}
+is the variance of $X$ when sampled according to the probability distribution $p$.
+By expanding
+\begin{align}
+  t_{\ell\mu} = \sqrt{\f{2\ell+1}{d}}\, \tau_{\ell\mu},
+  &&
+  \tau_{\ell\mu} \equiv \bk{s\mu;\ell,0|s\mu},
+\end{align}
+we can simplify
+\begin{align}
+  g_\ell\p{p} = \f{2\ell+1}{d} \times \sigma_p^2\p{\tau_\ell}.
+\end{align}
+We now seek to find an upper bound for $g_\ell\p{p}$, which will provide an upper bound for $\bbk{\epsilon_{i\ell}^2}$.
+To bound $g_\ell\p{p}$, we note that variance $\sigma_p^2\p{X}$ is maximal when $p$ has equal weight on the largest and smallest values of $X$, which implies that
+\begin{align}
+  \sigma_p^2\p{\tau_\ell} \le \gamma_\ell^2,
+  &&
+  \gamma_\ell \equiv \f{\max\tau_\ell - \min\tau_\ell}{2}.
+\end{align}
+Altogether, using additionally the fact that $g_0\p{p}=0$, we can bound
+\begin{align}
+  \bbk{\epsilon_{i\ell}^2}
+  = \f1n \times g_\ell\p{p^{(i)}}
+  \le \f{2\ell+1}{nd} \times \gamma_\ell^2
+  \times \p{1 - \delta_{\ell,0}}.
+  \label{eq:variance_bound}
+\end{align}
+Note that the bound in Eq.~\eqref{eq:variance_bound} is tight, as equality is achieved by the mixed state
+\begin{align}
+  \rho^\star
+  \equiv \f12 \p{\Pi_{\v v_i,\mu_{\t{max}}} + \Pi_{\v v_i,\mu_{\t{min}}}},
+\end{align}
+where $\mu_{\t{max}}$ ($\mu_{\t{min}}$) is the index that maximizes (minimizes) $\tau_{\ell\mu}$.
+In addition, the  Clebsch-Gordan coefficients $\tau_{\ell\mu}$ are at most $1$ in magnitude, which implies that $\gamma_\ell^2\le 1$.
+However, this bound on $\gamma_\ell^2$ is not tight in general, as we could have alternately bounded
 \begin{align}
   g_\ell\p{p} \le \sum_\mu t_{\ell\mu}^2 p_\mu
-  \le \max_\mu t_{\ell\mu}^2
-  = \f{2\ell+1}{d} \max_\mu \bk{s\mu;\ell,0|s\mu}^2
-  \le \f{2\ell+1}{d},
+  \le \sum_\mu t_{\ell\mu}^2 = 1,
 \end{align}
-which implies that
+which must also hold for the probability distribution $p^\star$ that maximizes $g_\ell$, i.e.
 \begin{align}
-  \bbk{\epsilon_{i\ell}^2} = \f1n g_\ell\p{p^{(i)}} \le \f{2\ell+1}{nd}.
-  \label{eq:variance_bound}
+  g_\ell\p{p^\star} = \f{2\ell+1}{d} \times \gamma_\ell^2 \le 1,
+  &&
+  \t{so}
+  &&
+  \gamma_\ell^2 \le \min\Set{1, \f{d}{2\ell+1}}.
 \end{align}
-Note that $t_{0,\mu}=1/\sqrt{d}$, so $f_0\p{p}=0$ and $\bbk{\epsilon_{i,0}^2}=0$.
 
@@
   \bbk{\epsilon_{i\ell}^2}
-  \le \f{2\ell+1}{nd} \p{1-\delta_{\ell,0}}.
+  \le \f{2\ell+1}{nd} \times \gamma_\ell^2 \times \p{1-\delta_{\ell,0}}.
 \end{align}
@@
   = \sum_{\ell,m} \Bbk{\abs{\tilde\rho_{\ell m}^V-\rho_{\ell m}^V}^2}
-  \le \f1{nd} \sum_{\ell>0} \p{2\ell+1} \S_{V\ell}^2,
+  \le \f1{nd} \sum_{\ell>0} \p{2\ell+1} \gamma_\ell^2 \S_{V\ell}^2,
 \end{align}

### 2020-11-08 sun_phases/spin_qudit_tomo.tex
moved around discussions of the tighter error bound

@@
 where $\set{\epsilon_j}$ are independent random variables and we use the double brackets $\bbk{\cdot}$ to denote statistical averaging.
-The assumption in Eq.~\eqref{eq:error_assumption} is reasonable when measurement error is dominated by experimental sources of noise, but breaks down when measurement error is limited by fundamental quantum (shot) noise.
-We relax the assumption of Eq.~\eqref{eq:error_assumption} and perform a proper treatment of measurement error in Appendix \ref{sec:error}.
-Our full treatment of measurement error yields identical conclusions, with an explicit form for $\epsilon$ in terms of the qudit dimension $d$ and the number of spin projection measurements made along every axis $\v v\in V$.
-To simplify our analysis, we therefore maintain the assumption of Eq.~\eqref{eq:error_assumption} for now, in which case the mean squared error with which $\tilde\rho^V_k$ approximates $\rho^V_k$ is
+In this case, the mean squared error with which $\tilde\rho^V_k$ approximates $\rho^V_k$ is
 \begin{align}
@@
   \le \epsilon^2 \S_V^2,
-  \label{eq:bound}
+  \label{eq:bound_eps}
 \end{align}
-where the {\it error scale} $\S_V$ of $V$, defined by
+where
 \begin{align}
-  \S_V \equiv \norm{M_V^{-1}} = \sqrt{\sum_k\f1{\p{M^V_k}^2}},
+  \S_V^2 \equiv \norm{M_V^{-1}}^2 = \sum_k\f1{\p{M^V_k}^2},
   \label{eq:scale}
 \end{align}
-is the (Hilbert-Schmidt, or Frobenius) norm of the right inverse $M_V^{-1}$ of $M_V$.
-Note that the error scale $\S_V$ diverges if $M_V$ is singular, in which case measuring spin projections along all axes in $V$ does not provide sufficient information to reconstruct arbitrary quantum states.
+is the squared (Hilbert-Schmidt, or Frobenius) norm of the right inverse $M_V^{-1}$ of $M_V$.
+Note that $\S_V^2$ diverges if $M_V$ is singular, in which case measuring spin projections along all axes in $V$ does not provide sufficient information to reconstruct arbitrary quantum states.
 
-The complexity of computing the error scale $\S_V$ is determined by the need to construct and compute the singular values of the measurement matrix $M_V$.
-This task can be greatly simplified by the fact that the degree $\ell$ of the transition operator $T_{\ell,0}$ is preserved under rotations, which implies that the unitary
+Computing $\S_V^2$ to certify the bound in Eq.~\eqref{eq:bound_eps} requires constructing the measurement matrix $M_V$ and finding its singular values.
+This complexity of this task can be greatly reduced by the fact that the degree $\ell$ of the transition operator $T_{\ell,0}$ is preserved under rotations, which implies that the unitary
 \begin{align}
@@
   &&
-  \S_{V\ell} \equiv \norm{M_{V\ell}^{-1}},
+  \S_{V\ell}^2 \equiv \norm{M_{V\ell}^{-1}}^2,
 \end{align}
 where $M_{V\ell}^{-1}$ is the right inverse of $M_{V\ell}$.
-Altogether, constructing the block $M_{V\ell}$ and computing its singular values takes at most $O(\abs{V}d^2)$ time (see Appendix \ref{sec:compute} for additional details on efficiently computing $\S_{V\ell}$ [\red{todo: write tutorial}]).
-If we assume that $\abs{V}\sim d$, then computing $\S_V$ to certify the bound in Eq.~\eqref{eq:bound} takes $O(d^4)$ serial or $O(d^3)$ parallel runtime (see Figure [\red{REF}]).
+Constructing the block $M_{V\ell}$ and computing its singular values takes at most $O(\abs{V}d^2)$ time (see Appendix \ref{sec:compute} for additional details on efficiently computing $\S_{V\ell}$ [\red{todo: write tutorial}]).
+Altogether, if we assume that $\abs{V}\sim d$, then computing $\S_V^2$ to certify the bound in Eq.~\eqref{eq:bound_eps} takes $O(d^4)$ serial or $O(d^3)$ parallel runtime (see Figure [\red{REF}]).
 [\red{todo: add figure showing runtimes as a function of $d$ with $\abs{V}=2d-1$; modify text appropriately after making the figure}].
@@
 
-[\red{todo: add a derivation of the error bound that assumes $n$ measurements of spin projection along each axis to Appendix \ref{sec:error}.
-  Briefly discuss these results in this section.}]
+The assumption that observables can be estimated up to uncorrelated noise with maximal variance $\epsilon^2$, summarized by Eq.~\eqref{eq:error_assumption}, is reasonable when measurement error is dominated by experimental sources of noise.
+However, this assumption breaks down when measurement error is limited by fundamental quantum (shot) noise.
+We relax the assumption of Eq.~\eqref{eq:error_assumption} and derive more fundamental error bounds in Appendix \ref{sec:error}.
+Therein, we find that if measurement error is dominated by shot noise, then the estimate $\tilde\rho_V$ built from $n$ independent measurements of spin projection along every axis $\v v\in V$ approximates the true state $\rho$ to within a mean squared distance
+\begin{align}
+  \Bbk{\norm{\tilde\rho_V-\rho}^2}
+  \le \epsilon_V^2
+  \equiv \f1{nd} \sum_{\ell>0} \p{2\ell+1} \gamma_\ell^2 \S_{V\ell},
+  \label{eq:bound}
+\end{align}
+where
+\begin{align}
+  \gamma_\ell
+  \equiv \f{\max_\mu\tau_{\ell\mu} - \min_\mu\tau_{\ell\mu}}{2}
+\end{align}
+is defined in terms of the Clebsch-Gordan coefficients
+\begin{align}
+  \tau_{\ell\mu} \equiv \bk{s\mu;\ell,0|s\mu}.
+\end{align}
+For reference, we note that $\epsilon_V^2<\S_V^2/n$.
 
@@
 
-The reconstruction error bound in Eq.~\eqref{eq:bound} provides a straightforward answer: the number of measurement axes should be chosen to minimize the product $\epsilon^2\S_V^2$, which bounds the expected error in the estimate $\tilde\rho_V$ of an unknown state $\rho$.
+The reconstruction error bound in Eq.~\eqref{eq:bound_eps} provides a straightforward answer: the number of measurement axes should be chosen to minimize the product $\epsilon^2\S_V^2$, which bounds the expected error in the estimate $\tilde\rho_V$ of an unknown state $\rho$.
 If there are a total of $N$ spin projection measurements onto $r$ quantization axes, then $n=N/r$ measurements are used to estimate spin projection onto any given axis.
@@
 In Section \ref{sec:bound} of the main text, we provided a reconstruction error bound using the assumption of Eq.~\eqref{eq:error_assumption}, namely that expectation values derived from spin projection measurements can be estimated up to uncorrelated errors with maximal variance $\epsilon^2$.
-This assumption is reasonable if measurement error is dominated by experimental sources of noise, and it yields a simple derivation of the reconstruction bound in Eq.~\eqref{eq:bound}.
+This assumption is reasonable if measurement error is dominated by experimental sources of noise, and it yields a simple derivation of the reconstruction bound in Eq.~\eqref{eq:bound_eps}.
 Nonetheless, there are two problems with the assumption of Eq.~\eqref{eq:error_assumption}:
@@
 \end{align}
-We now seek to find an upper bound for $g_\ell\p{p}$, which will provide an upper bound for $\bbk{\epsilon_{i\ell}^2}$.
-To bound $g_\ell\p{p}$, we note that variance $\sigma_p^2\p{X}$ is maximal when $p$ has equal weight on the largest and smallest values of $X$, which implies that
+We now seek to maximize $g_\ell\p{p}$, which will provide an upper bound for $\bbk{\epsilon_{i\ell}^2}$.
+In general, variance $\sigma_p^2\p{X}$ is maximal when $p$ has equal weight on the largest and smallest values of $X$, which implies that
 \begin{align}
@@
 In addition, the  Clebsch-Gordan coefficients $\tau_{\ell\mu}$ are at most $1$ in magnitude, which implies that $\gamma_\ell^2\le 1$.
-However, this bound on $\gamma_\ell^2$ is not tight in general, as we could have alternately bounded
+However, the bound on $\gamma_\ell^2\le1$ is not tight in general, as we could have alternately bounded
 \begin{align}
@@
 The block diagonal structure of this matrix implies that, rather than an integer index $k\in\set{1,2,\cdots,d^2}$, we can identify the singular values and right singular vectors of $UM_V$ (and equivalently $M_V$) by a block index $\ell\in\set{0,1,\cdots,d-1}$ and a sub-index $m=\set{-\ell,-\ell+1,\cdots,\ell}$.
-Making the replacement $k\to\p{\ell,m}$, we can thus repeat the derivations through Eq.~\eqref{eq:bound} of Section \ref{sec:bound} without the assumption of Eq.~\eqref{eq:error_assumption}, finding that
+Making the replacement $k\to\p{\ell,m}$, we can thus repeat the derivations through Eq.~\eqref{eq:bound_eps} of Section \ref{sec:bound} without the assumption of Eq.~\eqref{eq:error_assumption}, finding that
 \begin{align}
@@
   \S_{V\ell} \equiv \norm{M_{V\ell}^{-1}}
-  = \sqrt{\sum_m \p{M^V_{\ell m}}^{-2}}
+  = \sqrt{\sum_m \f1{\p{M^V_{\ell m}}^2}}
 \end{align}

### 2020-11-08 sun_phases/spin_qudit_tomo.tex
updated some discussions

@@
 \newcommand{\Q}{\mathcal{Q}}
-\newcommand{\R}{\mathcal{R}}
 \renewcommand{\S}{\mathcal{S}}
@@
   \S_V^2 \equiv \norm{M_V^{-1}}^2 = \sum_k\f1{\p{M^V_k}^2},
-  \label{eq:scale}
 \end{align}
@@
 where $M_{V\ell}^{-1}$ is the right inverse of $M_{V\ell}$.
-Constructing the block $M_{V\ell}$ and computing its singular values takes at most $O(\abs{V}d^2)$ time (see Appendix \ref{sec:compute} for additional details on efficiently computing $\S_{V\ell}$ [\red{todo: write tutorial}]).
+Constructing the block $M_{V\ell}$ and computing its singular values takes at most $O(\abs{V}d^2)$ time (see Appendix \ref{sec:compute} for additional details on efficiently computing $\S_{V\ell}^2$ [\red{todo: write tutorial}]).
 Altogether, if we assume that $\abs{V}\sim d$, then computing $\S_V^2$ to certify the bound in Eq.~\eqref{eq:bound_eps} takes $O(d^4)$ serial or $O(d^3)$ parallel runtime (see Figure [\red{REF}]).
@@
 We relax the assumption of Eq.~\eqref{eq:error_assumption} and derive more fundamental error bounds in Appendix \ref{sec:error}.
-Therein, we find that if measurement error is dominated by shot noise, then the estimate $\tilde\rho_V$ built from $n$ independent measurements of spin projection along every axis $\v v\in V$ approximates the true state $\rho$ to within a mean squared distance
+Therein, we find that if measurement error is dominated by shot noise, then the estimate $\tilde\rho_V$ built from $n$ independent measurements of spin projection along every axis $\v v\in V$ approximates the true state $\rho$ to within a mean squared distance $\epsilon_V^2/n$, i.e.
 \begin{align}
   \Bbk{\norm{\tilde\rho_V-\rho}^2}
-  \le \epsilon_V^2
-  \equiv \f1{nd} \sum_{\ell>0} \p{2\ell+1} \gamma_\ell^2 \S_{V\ell},
+  \le \f{\epsilon_V^2}{n},
   \label{eq:bound}
 \end{align}
-where
+where the {\it error scale} $\epsilon_V^2$ is
+\begin{align}
+  \epsilon_V^2 \equiv
+  \f1{nd} \sum_{\ell>0} \p{2\ell+1} \gamma_\ell^2 \S_{V\ell}^2,
+  \label{eq:scale}
+\end{align}
+and $\gamma_\ell$ is defined using Clebsch-Gordan coefficients:
 \begin{align}
   \gamma_\ell
-  \equiv \f{\max_\mu\tau_{\ell\mu} - \min_\mu\tau_{\ell\mu}}{2}
-\end{align}
-is defined in terms of the Clebsch-Gordan coefficients
-\begin{align}
+  \equiv \f{\max_\mu\tau_{\ell\mu} - \min_\mu\tau_{\ell\mu}}{2},
+  &&
   \tau_{\ell\mu} \equiv \bk{s\mu;\ell,0|s\mu}.
 \end{align}
-For reference, we note that $\epsilon_V^2<\S_V^2/n$.
+For reference, we note that the error scale $\epsilon_V^2<\S_V^2$.
 
@@
 
-The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and optimize (using any standard minimization algorithm) the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the error scale $\S_V$ defined in Eq.~\eqref{eq:scale}.
-If $\abs{V}$ is too large for optimization to be practical, one can instead generate many random sets of measurement axes, and then choose the set with the smallest error scale $\S_V$.
-After choosing a set of measurement axes $V$, experimentally characterize expectation values of spin projections, $\bk{\Pi_{\v v\mu}}_\rho$, which determine the expectation values $\bk{T_{\v v\ell,0}}_\rho$.
-If the experimental estimates of $\bk{T_{\v v\ell,0}}_\rho$ are accurate to within an error $\lesssim\epsilon$, then the state $\tilde\rho_V$ defined in Eq.~\eqref{eq:reconstructed_state} is an approximation of the true state $\rho$ to within a (Euclidean) distance $\lesssim\epsilon\S_V$.
-Furthermore, if the empirical estimate $\tilde\rho_V$ has negative eigenvalues, its distance from $\rho$ can be further reduced with maximum-likelihood corrections \cite{smolin2012efficient}.
+The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and optimize (using any standard minimization algorithm) the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the error scale $\epsilon_V^2$ defined in Eq.~\eqref{eq:scale}.
+If $\abs{V}$ is too large for such optimization  to be practical, one can instead generate many random sets of measurement axes, and then choose the set with the smallest error scale $\epsilon_V^2$.
+After choosing a set of measurement axes $V$, experimentally characterize the expectation values $\bk{\Pi_{\v v\mu}}_\rho$ by making measuring spin projection along every axis $\v v\in V$.
+These spin projections determine the expectation values $\bk{T_{\v v\ell,0}}_\rho$, whose estimates can be used to construct the state $\tilde\rho$ defined in Eq.~\eqref{eq:reconstructed_state}.
+If spin projection was measured $n$ times along every axis, then the reconstructed state $\tilde\rho$ approximates the true state $\rho$ to within a mean squared (Euclidean) distance $\epsilon_V^2/n$.
+Furthermore, if $\tilde\rho_V$ has negative eigenvalues, its distance from $\rho$ can be further reduced with maximum-likelihood corrections \cite{smolin2012efficient}.
 
@@
 
-As a last point, we mention that the error scale $\S_V$ can also evaluate the ``quality'' of measurement axes $V$ chosen with existing methods for sparse signal recovery on the sphere \cite{rauhut2011sparse, alem2012sparse}.
+As a last point, we mention that the error scale $\epsilon_V^2$ can also be used to evaluate the ``quality'' of measurement axes $V$ chosen with existing methods for sparse signal recovery on the sphere \cite{rauhut2011sparse, alem2012sparse}.
 In practice, we find that uniform random sampling outperforms these methods, in the sense that it generates low-error-scale sets of measurement axes with higher probability.
@@
 Even so, $2d-1$ measurement axes may not be the ``optimal'' number to use.
-Increasing the number of measurement axes generally decreases the error scale $\S_V$, but at the cost of having to estimate more observables.
-If we fix the total number of measurements (or experimental trials), then having to estimate more observables means that fewer measurements are devoted to estimating any particular observable, which in turn increases the error $\epsilon$ on those estimates.
+Increasing the number of measurement axes generally decreases the error scale $\epsilon_V^2$, but at the cost of having to estimate more observables.
+At a fixed total number of measurements, $N=n\abs{V}$, increasing the number of measurement axes $\abs{V}$ reduces the number of measurements $n$ along each axis $\v v\in V$.
 This trade-off begs the question: how many measurement axes should be chosen to perform spin qudit tomography?
 
-The reconstruction error bound in Eq.~\eqref{eq:bound_eps} provides a straightforward answer: the number of measurement axes should be chosen to minimize the product $\epsilon^2\S_V^2$, which bounds the expected error in the estimate $\tilde\rho_V$ of an unknown state $\rho$.
-If there are a total of $N$ spin projection measurements onto $r$ quantization axes, then $n=N/r$ measurements are used to estimate spin projection onto any given axis.
-The variance in estimates of spin projection is then $\epsilon^2\sim1/n\sim r$ at a fixed total number of measurements $N$.
-Altogether, the number of measurement axes should be chosen in such a way as to minimize $r\S_V^2$ with $\abs{V}=r$.
+The reconstruction error bound in Eq.~\eqref{eq:bound} nominally provides a straightforward answer: at a fixed total number of measurements $N=n\abs{V}$, the number of measurement axes should be chosen to minimize $\epsilon_V^2/n\propto\epsilon_V^2\abs{V}$.
 
@@
 In addition, the  Clebsch-Gordan coefficients $\tau_{\ell\mu}$ are at most $1$ in magnitude, which implies that $\gamma_\ell^2\le 1$.
-However, the bound on $\gamma_\ell^2\le1$ is not tight in general, as we could have alternately bounded
+However, the bound on $\gamma_\ell^2\le1$ is not tight in general, as we could have alternatively bounded
 \begin{align}
@@
 \begin{align}
-  \S_{V\ell} \equiv \norm{M_{V\ell}^{-1}}
-  = \sqrt{\sum_m \f1{\p{M^V_{\ell m}}^2}}
+  \S_{V\ell}^2 \equiv \norm{M_{V\ell}^{-1}}^2
+  = \sum_m \f1{\p{M^V_{\ell m}}^2}
 \end{align}
-is the (Hilbert-Schmidt, or Frobenius) norm of the right inverse of $M_{V\ell}$.
+is the squared (Hilbert-Schmidt, or Frobenius) norm of the right inverse of $M_{V\ell}$.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Computing the error scale}
+\section{Computing singular values of the measurement matrix}
 \label{sec:compute}

### 2020-11-08 sun_phases/spin_qudit_tomo.tex
a bit more tightening of the analytical error bound

@@
 
+\newcommand{\ZZ}{\mathbb{Z}}
+
 \def\obra#1{\mathinner{({#1}|}}
@@
 \newcommand{\red}[1]{{\color{red} #1}}
-\newcommand{\todo}[1]{{\color{blue} To do: #1}}
 
@@
   \epsilon_V^2 \equiv
-  \f1{nd} \sum_{\ell>0} \p{2\ell+1} \gamma_\ell^2 \S_{V\ell}^2,
+  \f1d \sum_{\ell>0} \p{2\ell+1} \gamma_\ell^2 \S_{V\ell}^2,
   \label{eq:scale}
@@
 \end{align}
-where $\mu_{\t{max}}$ ($\mu_{\t{min}}$) is the index that maximizes (minimizes) $\tau_{\ell\mu}$.
-In addition, the  Clebsch-Gordan coefficients $\tau_{\ell\mu}$ are at most $1$ in magnitude, which implies that $\gamma_\ell^2\le 1$.
-However, the bound on $\gamma_\ell^2\le1$ is not tight in general, as we could have alternatively bounded
+where $\mu_{\t{max}}$ ($\mu_{\t{min}}$) is any index that maximizes (minimizes) $\tau_{\ell\mu}$.
+
+Though the result in Eq.~\eqref{eq:variance_bound} is sufficient to proceed with our derivation of an improved tomographic error bound, we first take some time to better understand the bound on $\bbk{\epsilon_{i\ell}^2}$ analytically.
+To this end, we first note that the Clebsch-Gordan coefficients $\tau_{\ell\mu}$ are at most $1$ in magnitude, which implies that $\gamma_\ell^2\le 1$.
+Moreover, we can also bound
 \begin{align}
@@
   \le \sum_\mu t_{\ell\mu}^2 = 1,
-\end{align}
-which must also hold for the probability distribution $p^\star$ that maximizes $g_\ell$, i.e.
-\begin{align}
-  g_\ell\p{p^\star} = \f{2\ell+1}{d} \times \gamma_\ell^2 \le 1,
   &&
@@
   &&
-  \gamma_\ell^2 \le \min\Set{1, \f{d}{2\ell+1}}.
+  \bbk{\epsilon_{i\ell}^2} \le \f1n \times \min\Set{\f{2\ell+1}{d}, 1}.
+\end{align}
+If $d$ is even, then we can get a tighter analytical bound on $\bbk{\epsilon_{i\ell}^2}$ by considering
+\begin{align}
+  g_\ell\p{p} \le \sum_\mu t_{\ell\mu}^2 p_\mu \le t_{\ell\mu_{\t{max}}}^2,
+\end{align}
+where $\mu_{\t{max}}$ is any index maximizing $t_{\ell\mu}$.
+The symmetries of the Clebsch-Gordan coefficients imply that $t_{\ell\mu_{\t{max}}}^2 = t_{\ell,-\mu_{\t{max}}}^2$, so if $\mu_{\t{max}} \ne0$ (which is always true when $d$ is even because $s,\mu\notin\ZZ$), then
+\begin{align}
+  g_\ell\p{p} \le t_{\ell\mu_{\t{max}}}^2
+  = \f12 \p{t_{\ell\mu_{\t{max}}}^2 + t_{\ell,-\mu_{\t{max}}}^2}
+  \le \f12 \sum_\mu t_{\ell\mu}^2
+  = \f12,
+  &&
+  \t{so}
+  &&
+  \bbk{\epsilon_{i\ell}^2}
+  \le \f1n \times \min\Set{\f{2\ell+1}{d}, \f12}.
 \end{align}

### 2020-11-09 sun_phases/spin_qudit_tomo.tex
improved simplified 'analytical' form of error bound

@@
 \end{align}
-If $d$ is even, then we can get a tighter analytical bound on $\bbk{\epsilon_{i\ell}^2}$ by considering
+We can get a tighter analytical bound on $\bbk{\epsilon_{i\ell}^2}$ by considering that if $\mu_{\t{max}}\ne0$, then
 \begin{align}
-  g_\ell\p{p} \le \sum_\mu t_{\ell\mu}^2 p_\mu \le t_{\ell\mu_{\t{max}}}^2,
+  g_\ell\p{p} \le \sum_\mu t_{\ell\mu}^2 p_\mu
+  \le t_{\ell\mu_{\t{max}}}^2
+  = \f12 \p{t_{\ell\mu_{\t{max}}}^2 + t_{\ell,-\mu_{\t{max}}}^2}
+  \stackrel{\mu_{\t{max}}\ne0}{\le} \f12 \sum_\mu t_{\ell\mu}^2
+  = \f12,
 \end{align}
-where $\mu_{\t{max}}$ is any index maximizing $t_{\ell\mu}$.
-The symmetries of the Clebsch-Gordan coefficients imply that $t_{\ell\mu_{\t{max}}}^2 = t_{\ell,-\mu_{\t{max}}}^2$, so if $\mu_{\t{max}} \ne0$ (which is always true when $d$ is even because $s,\mu\notin\ZZ$), then
+where we used the fact that $t_{\ell\mu_{\t{max}}}^2 = t_{\ell,-\mu_{\t{max}}}^2$ due to the symmetries of the Clebsch-Gordan coefficients.
+If $\mu_{\t{max}}=0$, then
 \begin{align}
-  g_\ell\p{p} \le t_{\ell\mu_{\t{max}}}^2
-  = \f12 \p{t_{\ell\mu_{\t{max}}}^2 + t_{\ell,-\mu_{\t{max}}}^2}
-  \le \f12 \sum_\mu t_{\ell\mu}^2
-  = \f12,
+  t_{\ell\mu_{\t{max}}}^2 + 2 t_{\ell\mu_{\t{min}}}^2
+  = t_{\ell\mu_{\t{max}}}^2 + t_{\ell\mu_{\t{min}}}^2
+  + t_{\ell,-\mu_{\t{min}}}^2
+  \stackrel{\mu_{\t{max}}=0}{\le}
+  \sum_\mu t_{\ell\mu}^2 = 1,
   &&
@@
   &&
+  \abs{t_{\ell\mu_{\t{min}}}}
+  \stackrel{\mu_{\t{max}}=0}{\le}
+  \sqrt{\f{1 - t_{\ell\mu_{\t{max}}}^2}{2}}.
+\end{align}
+This inequality lets us bound
+\begin{align}
+  \sqrt{\f{2\ell+1}{d}}\, \gamma_\ell
+  = \f12 \p{t_{\ell\mu_{\t{max}}} - t_{\ell\mu_{\t{min}}}}
+  \le \f12 \p{t_{\ell\mu_{\t{max}}} + \abs{t_{\ell\mu_{\t{min}}}}}
+  \stackrel{\mu_{\t{max}}=0}{\le}
+  \f12 t_{\ell\mu_{\t{max}}}
+  + \f12\sqrt{\f{1 - t_{\ell\mu_{\t{max}}}^2}{2}}
+  \equiv \Gamma\p{t_{\ell\mu_{\t{max}}}}.
+\end{align}
+If is straightforward to show that $\Gamma\p{M}$ has a single maximum at $M=\sqrt{2/3}$, and that $\Gamma^\star\equiv\Gamma\p{\sqrt{2/3}}=\sqrt{3/8}$, so
+\begin{align}
+  g_\ell\p{p} \le \f{2\ell+1}{d}\, \gamma_\ell^2
+  \stackrel{\mu_{\t{max}}=0}{\le} \p{\Gamma^\star}^2
+  = \f38 < \f12.
+\end{align}
+Altogether, we thus find that in all cases
+\begin{align}
   \bbk{\epsilon_{i\ell}^2}

### 2020-11-09 sun_phases/spin_qudit_tomo.tex
improved simplified 'analytical' form of error bound
comment on tightness (or rather, looseness) of the error bound

@@
 \begin{align}
-  \Bbk{\norm{\tilde\rho_V-\rho}^2}
-  \le \f{\epsilon_V^2}{n},
+  \Bbk{\norm{\tilde\rho_V-\rho}^2} < \f{\epsilon_V^2}{n},
   \label{eq:bound}
@@
 \end{align}
-For reference, we note that the error scale $\epsilon_V^2<\S_V^2$.
+For comparison, we note that the error scale $\epsilon_V^2<\S_V^2/2$.
 
@@
   \bbk{\epsilon_{i\ell}^2}
-  \le \f{2\ell+1}{nd} \times \gamma_\ell^2 \times \p{1-\delta_{\ell,0}}.
+  < \f{2\ell+1}{nd} \times \gamma_\ell^2
+  \times \p{1-\delta_{\ell,0}},
+  \label{eq:bound_single}
 \end{align}
+where $i$ indexes an axis $\v v_i\in V$.
 Altogether, we thus find that $n$ measurements of spin projection along every axis $\v v\in V$ allows us to construct an empirical estimate $\tilde\rho_V$ of an unknown state $\rho$ that is accurate to within a mean squared distance
@@
   = \sum_{\ell,m} \Bbk{\abs{\tilde\rho_{\ell m}^V-\rho_{\ell m}^V}^2}
-  \le \f1{nd} \sum_{\ell>0} \p{2\ell+1} \gamma_\ell^2 \S_{V\ell}^2,
+  < \f1{nd} \sum_{\ell>0} \p{2\ell+1} \gamma_\ell^2 \S_{V\ell}^2.
+  \label{eq:bound_full}
 \end{align}
-where
-\begin{align}
-  \S_{V\ell}^2 \equiv \norm{M_{V\ell}^{-1}}^2
-  = \sum_m \f1{\p{M^V_{\ell m}}^2}
-\end{align}
-is the squared (Hilbert-Schmidt, or Frobenius) norm of the right inverse of $M_{V\ell}$.
+We note that whereas the bound on $\bbk{\epsilon_{i\ell}^2}$ in Eq.~\eqref{eq:variance_bound} is tight, the bounds in Eq.~\eqref{eq:bound_single} and Eq.~\eqref{eq:bound_full} are not: the upper bound on $\bbk{\epsilon_{i\ell}^2}$ cannot be achieved simultaneously for all axes $\v v_i$ or degrees $\ell$.
+In principle, there is therefore room for improvement on the bound in Eq.~\eqref{eq:bound_full}.
 

### 2020-11-09 sun_phases/spin_qudit_tomo.tex
cleaned up derivation of bounds on individual variances

@@
 For convenience, we index the tensor factors of $\rho^{\otimes N}$ by the integers $\p{i,j}$, with $i\in\set{1,2,\cdots,\abs{V}}$ specifying a measurement axis $\v v_i\in V$ and $j\in\set{1,2,\cdots,n}$ specifying the copy of $\rho$ prepared for the $j$-th measurement spin projection along a particular axis.
-We then define the projectors $\Pi_{i\mu}\equiv\op{\mu_{\v v_i}}$ onto states $\ket{\mu_{\v v_i}}$ with definite spin projection $\mu$ onto axis $\v v_i\in V$, and define $\Pi_{i\mu}^{(j)}$ to be an $N$-qudit operator with $\Pi_{i\mu}$ on the $\p{i,j}$-th tensor factor and the identity elsewhere.
-We denote the experimental outcome of measuring $\Pi_{i\mu}$ in the $\p{i,j}$-th copy of $\rho$ by $\tilde \Pi_{i\mu}^{(j)}\in\set{0,1}$.
-In other words, $\tilde \Pi_{i\mu}^{(j)}$ is the ``single-shot estimate'' of $\Pi_{i\mu}$, with $\tilde \Pi_{i\mu}^{(j)}=1$ if outcome $\mu$ was observed on the $\p{i,j}$-th experimental trial, and $\tilde \Pi_{i\mu}^{(j)}=0$ otherwise.
+We then define the projectors $\Pi_{i\mu}\equiv\op{\mu_{\v v_i}}$ onto states $\ket{\mu_{\v v_i}}$ with definite spin projection $\mu$ onto axis $\v v_i\in V$, and define $\Pi_{i\mu}^j$ to be an $N$-qudit operator with $\Pi_{i\mu}$ on the $\p{i,j}$-th tensor factor and the identity elsewhere.
+We denote the experimental outcome of measuring $\Pi_{i\mu}$ in the $\p{i,j}$-th copy of $\rho$ by $\tilde \Pi_{i\mu}^j\in\set{0,1}$.
+In other words, $\tilde \Pi_{i\mu}^j$ is the ``single-shot estimate'' of $\Pi_{i\mu}$, with $\tilde \Pi_{i\mu}^j=1$ if outcome $\mu$ was observed on the $\p{i,j}$-th experimental trial, and $\tilde \Pi_{i\mu}^j=0$ otherwise.
 An empirical estimate of the expectation value $\bk{\Pi_{i\mu}}_\rho$ is then provided by the fraction of times that outcome $\mu$ was observed when measuring spin projection along axis $\v v_i$, that is
 \begin{align}
-  \tilde \Pi_{i\mu} \equiv \f1n \sum_{j=1}^n \tilde \Pi_{i\mu}^{(j)}
-  \approx \f1n \sum_{j=1}^n \tr\p{\Pi_{i\mu}^{(j)} \rho^{\otimes N}}
+  \tilde \Pi_{i\mu} \equiv \f1n \sum_{j=1}^n \tilde \Pi_{i\mu}^j
+  \approx \f1n \sum_{j=1}^n \tr\p{\Pi_{i\mu}^j \rho^{\otimes N}}
   = \tr\p{\Pi_{i\mu} \rho}.
@@
 \begin{align}
-  \bar \Pi_{i\mu} \equiv \f1n \sum_{j=1}^n \Pi_{i\mu}^{(j)}
+  \bar \Pi_{i\mu} \equiv \f1n \sum_{j=1}^n \Pi_{i\mu}^j
   \label{eq:mean_proj}
@@
   = \f1{n^2} \sum_{j,j'=1}^n
-  \sp{\bk{\Pi_{i\mu}^{(j)} \Pi_{i'\mu'}^{(j')}}_{\rho^{\otimes N}}
-    - \bk{\Pi_{i\mu}^{(j)}}_{\rho^{\otimes N}}
-    \bk{\Pi_{i'\mu'}^{(j')}}_{\rho^{\otimes N}}}.
+  \sp{\bk{\Pi_{i\mu}^j \Pi_{i'\mu'}^{j'}}_{\rho^{\otimes N}}
+    - \bk{\Pi_{i\mu}^j}_{\rho^{\otimes N}}
+    \bk{\Pi_{i'\mu'}^{j'}}_{\rho^{\otimes N}}}.
   \label{eq:proj_cov_start}
 \end{align}
-If $\p{i,j}\ne\p{i',j'}$, then $\Pi_{i\mu}^{(j)}$ and $\Pi_{i'\mu'}^{(j')}$ address different tensor factors of the product state $\rho^{\otimes N}$, so the expectation value of their product factorizes due to the identity $\tr\sp{\p{A\otimes B}\p{A'\otimes B'}} = \tr\p{AA'}\times\tr\p{BB'}$.
+If $\p{i,j}\ne\p{i',j'}$, then $\Pi_{i\mu}^j$ and $\Pi_{i'\mu'}^{j'}$ address different tensor factors of the product state $\rho^{\otimes N}$, so the expectation value of their product factorizes due to the identity $\tr\sp{\p{A\otimes B}\p{A'\otimes B'}} = \tr\p{AA'}\times\tr\p{BB'}$.
 The terms in Eq.~\eqref{eq:proj_cov_start} with $\p{i,j}\ne\p{i',j'}$ therefore vanish, so
@@
   \bbk{\epsilon_{i\mu} \epsilon_{i'\mu'}}
-  &= \delta_{ii'} \times \f1{n^2} \sum_{j=1}^n \sp{\bk{\Pi_{i\mu}^{(j)} \Pi_{i\mu'}^{(j)}}_{\rho^{\otimes N}}
-    - \bk{\Pi_{i\mu}^{(j)}}_{\rho^{\otimes N}}
-    \bk{\Pi_{i\mu'}^{(j)}}_{\rho^{\otimes N}}} \\
+  &= \delta_{ii'} \times \f1{n^2} \sum_{j=1}^n \sp{\bk{\Pi_{i\mu}^j \Pi_{i\mu'}^j}_{\rho^{\otimes N}}
+    - \bk{\Pi_{i\mu}^j}_{\rho^{\otimes N}}
+    \bk{\Pi_{i\mu'}^j}_{\rho^{\otimes N}}} \\
   &= \delta_{ii'} \times \f1n \sp{\bk{\Pi_{i\mu} \Pi_{i\mu'}}_\rho
     - \bk{\Pi_{i\mu}}_\rho \bk{\Pi_{i\mu'}}_\rho} \\
-  &= \delta_{ii'} \times \f1n \sp{\delta_{\mu\mu'} p_\mu^{(i)} - p_\mu^{(i)} p_{\mu'}^{(i)}},
+  &= \delta_{ii'} \times \f1n \sp{\delta_{\mu\mu'} p_\mu^i - p_\mu^i p_{\mu'}^i},
 \end{align}
-where $p_\mu^{(i)} \equiv \bk{\Pi_{i\mu}}_\rho = \bk{\mu_{\v v_i}|\rho|\mu_{\v v_i}}$ are the diagonal entries of $\rho$ in the basis of spin projections along axis $\v v_i$.
+where $p_\mu^i \equiv \bk{\Pi_{i\mu}}_\rho = \bk{\mu_{\v v_i}|\rho|\mu_{\v v_i}}$ are the diagonal entries of $\rho$ in the basis of spin projections along axis $\v v_i$.
 
@@
   = \delta_{ii'} \times \f1n \sum_{\mu,\mu'} t_{\ell\mu} t_{\ell'\mu'}
-  \sp{\delta_{\mu\mu'} p_\mu^{(i)} - p_\mu^{(i)} p_{\mu'}^{(i)}}.
+  \sp{\delta_{\mu\mu'} p_\mu^i - p_\mu^i p_{\mu'}^i}.
 \end{align}
-In order to establish an upper bound on reconstruction error for an unknown quantum state, we need an upper bound on the variance $\bbk{\epsilon_{i\ell}^2}$ in particular.
-To this end, for any multinomial probability distribution $p\equiv\p{p_1,p_2,\cdots,p_d}$ we define the function
+In order to establish an upper bound on reconstruction error for an unknown quantum state, we will need an upper bound on the variance $\bbk{\epsilon_{i\ell}^2}$ in particular.
+To this end, we observe that
 \begin{align}
-  g_\ell\p{p} \equiv \sum_{\mu,\mu'} t_{\ell\mu} t_{\ell\mu'}
-  \p{\delta_{\mu\mu'} p_\mu - p_\mu p_{\mu'}}
-  = \sum_\mu t_{\ell\mu}^2 p_\mu - \p{\sum_\mu t_{\ell\mu} p_\mu}^2
-  = \sigma_p^2\p{t_\ell},
+  \bbk{\epsilon_{i\ell}^2} = \f1n \times \sigma_{p^i}^2\p{t_\ell},
 \end{align}
@@
   \sigma_p^2\p{X}
-  \equiv \sum_\mu p_\mu \p{X_\mu - \sum_\nu p_\nu X_\nu}^2
-  = \sum_\mu p_\mu X_\mu^2 - \p{\sum_\mu p_\mu X_\mu}^2
+  \equiv \sum_\mu p_\mu X_\mu^2 - \p{\sum_\mu p_\mu X_\mu}^2
+  = \sum_\mu p_\mu \p{X_\mu - \sum_\nu p_\nu X_\nu}^2
 \end{align}
 is the variance of $X$ when sampled according to the probability distribution $p$.
-By expanding
+The variance $\sigma_p^2\p{X}$ is maximal when $p$ has equal weight on the largest and smallest values of $X$, which implies that
 \begin{align}
-  t_{\ell\mu} = \sqrt{\f{2\ell+1}{d}}\, \tau_{\ell\mu},
+  \sigma_p^2\p{t_\ell} \le \Gamma_\ell^2,
   &&
-  \tau_{\ell\mu} \equiv \bk{s\mu;\ell,0|s\mu},
+  \Gamma_\ell \equiv \f{\max_\mu t_{\ell\mu} - \min_\mu t_{\ell\mu}}{2}.
 \end{align}
-we can simplify
-\begin{align}
-  g_\ell\p{p} = \f{2\ell+1}{d} \times \sigma_p^2\p{\tau_\ell}.
-\end{align}
-We now seek to maximize $g_\ell\p{p}$, which will provide an upper bound for $\bbk{\epsilon_{i\ell}^2}$.
-In general, variance $\sigma_p^2\p{X}$ is maximal when $p$ has equal weight on the largest and smallest values of $X$, which implies that
-\begin{align}
-  \sigma_p^2\p{\tau_\ell} \le \gamma_\ell^2,
-  &&
-  \gamma_\ell \equiv \f{\max\tau_\ell - \min\tau_\ell}{2}.
-\end{align}
-Altogether, using additionally the fact that $g_0\p{p}=0$, we can bound
+Altogether, using additionally the fact that $\sigma_p^2\p{t_0}=0$ (because all $t_{0,\mu}=1/\sqrt{d}$) we can bound
 \begin{align}
   \bbk{\epsilon_{i\ell}^2}
-  = \f1n \times g_\ell\p{p^{(i)}}
-  \le \f{2\ell+1}{nd} \times \gamma_\ell^2
-  \times \p{1 - \delta_{\ell,0}}.
+  \le \f1n \times \Gamma_\ell^2 \times \p{1 - \delta_{\ell,0}}
+  = \f{2\ell+1}{nd} \times \gamma_\ell^2
+  \times \p{1 - \delta_{\ell,0}},
   \label{eq:variance_bound}
+\end{align}
+where $\gamma_\ell$ is defined using Clebsch-Gordan coefficients:
+\begin{align}
+  \gamma_\ell \equiv \f{\max_\mu\tau_{\ell\mu} - \min_\mu\tau_{\ell\mu}}{2},
+  &&
+  \tau_{\ell\mu} \equiv \bk{s\mu;\ell,0|s\mu}.
 \end{align}
@@
 
-Though the result in Eq.~\eqref{eq:variance_bound} is sufficient to proceed with our derivation of an improved tomographic error bound, we first take some time to better understand the bound on $\bbk{\epsilon_{i\ell}^2}$ analytically.
+Though the result in Eq.~\eqref{eq:variance_bound} is sufficient to proceed with our derivation of an improved tomographic reconstruction error bound, we first take some time to better understand the bound on $\bbk{\epsilon_{i\ell}^2}$ analytically.
 To this end, we first note that the Clebsch-Gordan coefficients $\tau_{\ell\mu}$ are at most $1$ in magnitude, which implies that $\gamma_\ell^2\le 1$.
@@
 \begin{align}
-  g_\ell\p{p} \le \sum_\mu t_{\ell\mu}^2 p_\mu
+  \sigma_p^2\p{t_\ell} \le \sum_\mu t_{\ell\mu}^2 p_\mu
   \le \sum_\mu t_{\ell\mu}^2 = 1,
@@
 \end{align}
-We can get a tighter analytical bound on $\bbk{\epsilon_{i\ell}^2}$ by considering that if $\mu_{\t{max}}\ne0$, then
+We can get a tighter analytical bound on $\bbk{\epsilon_{i\ell}^2}$ by considering the fact that $t_{\ell\mu_{\t{max}}}^2 = t_{\ell,-\mu_{\t{max}}}^2$ due to the symmetries of the Clebsch-Gordan coefficients, so if $\mu_{\t{max}}\ne0$, then
 \begin{align}
-  g_\ell\p{p} \le \sum_\mu t_{\ell\mu}^2 p_\mu
+  \sigma_p^2\p{t_\ell} \le \sum_\mu t_{\ell\mu}^2 p_\mu
   \le t_{\ell\mu_{\t{max}}}^2
@@
   \stackrel{\mu_{\t{max}}\ne0}{\le} \f12 \sum_\mu t_{\ell\mu}^2
-  = \f12,
+  = \f12.
 \end{align}
-where we used the fact that $t_{\ell\mu_{\t{max}}}^2 = t_{\ell,-\mu_{\t{max}}}^2$ due to the symmetries of the Clebsch-Gordan coefficients.
-If $\mu_{\t{max}}=0$, then
+If $\mu_{\t{max}}=0$, then similarly
 \begin{align}
@@
 \begin{align}
-  \sqrt{\f{2\ell+1}{d}}\, \gamma_\ell
+  \Gamma_\ell
   = \f12 \p{t_{\ell\mu_{\t{max}}} - t_{\ell\mu_{\t{min}}}}
@@
   + \f12\sqrt{\f{1 - t_{\ell\mu_{\t{max}}}^2}{2}}
-  \equiv \Gamma\p{t_{\ell\mu_{\t{max}}}}.
+  \equiv g\p{t_{\ell\mu_{\t{max}}}}.
 \end{align}
-If is straightforward to show that $\Gamma\p{M}$ has a single maximum at $M=\sqrt{2/3}$, and that $\Gamma^\star\equiv\Gamma\p{\sqrt{2/3}}=\sqrt{3/8}$, so
+If is straightforward to show that $g\p{M}$ has a single maximum at $M=\sqrt{2/3}$, and that $g^\star\equiv g\p{\sqrt{2/3}}=\sqrt{3/8}$, so
 \begin{align}
-  g_\ell\p{p} \le \f{2\ell+1}{d}\, \gamma_\ell^2
-  \stackrel{\mu_{\t{max}}=0}{\le} \p{\Gamma^\star}^2
+  \sigma_p^2\p{t_\ell} \le \Gamma_\ell^2
+  \stackrel{\mu_{\t{max}}=0}{\le} \p{g^\star}^2
   = \f38 < \f12.

### 2020-11-09 sun_phases/spin_qudit_tomo.tex
removed one minor comment and added another

@@
 [\red{todo: add figure showing runtimes as a function of $d$ with $\abs{V}=2d-1$; modify text appropriately after making the figure}].
-For comparison, we note that existing algorithms for signal recovery on the sphere \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality} similarly have $O(d^3)$ parallel runtime with $\abs{V}\gtrsim d$.
 
@@
 \end{align}
+If $d$ even or $\ell$ is odd, then $\gamma_\ell=\max_\mu\tau_{\ell\mu}$.
 For comparison, we note that the error scale $\epsilon_V^2<\S_V^2/2$.

### 2020-11-09 sun_phases/spin_qudit_tomo.tex
added a figure of serial runtime, and tweaked some discussions

@@
 Constructing the block $M_{V\ell}$ and computing its singular values takes at most $O(\abs{V}d^2)$ time (see Appendix \ref{sec:compute} for additional details on efficiently computing $\S_{V\ell}^2$ [\red{todo: write tutorial}]).
-Altogether, if we assume that $\abs{V}\sim d$, then computing $\S_V^2$ to certify the bound in Eq.~\eqref{eq:bound_eps} takes $O(d^4)$ serial or $O(d^3)$ parallel runtime (see Figure [\red{REF}]).
-[\red{todo: add figure showing runtimes as a function of $d$ with $\abs{V}=2d-1$; modify text appropriately after making the figure}].
+Altogether, if we assume that $\abs{V}\sim d$, then computing $\S_V^2$ to certify the bound in Eq.~\eqref{eq:bound_eps} takes $O(d^4)$ serial or $O(d^3)$ parallel runtime (see Figure \ref{fig:times}).
+
+\begin{figure}
+  \centering
+  \includegraphics{error_scale/error_times.pdf}
+  \caption{Serial runtime to compute $\S_V$ with $\abs{V}=2d-1$ randomly chosen measurement axes, averaged over 100 calculations of $\S_V$ or 5 minutes of runtime (for each $d$), whichever comes first.
+    Dashed line shows a fit to runtime $t=c d^\alpha$ for all $d\ge100$, finding $\alpha\approx 3.7$.
+    These results do not include the $O(d^3)$ runtime for pre-computing $\pi/2$-pulses $e^{\i\p{\pi/2}S_\y}$ (see Appendix \ref{sec:compute}), which can be recycled for every new choice of measurement axes $V$.}
+  \label{fig:times}
+\end{figure}
 
@@
 If $d$ even or $\ell$ is odd, then $\gamma_\ell=\max_\mu\tau_{\ell\mu}$.
-For comparison, we note that the error scale $\epsilon_V^2<\S_V^2/2$.
+For comparison with the previous bound in Eq.~\eqref{eq:bound_eps}, we note that the error scale $\epsilon_V^2<\S_V^2/2$, so the previous bound still holds with the replacement $\epsilon^2\to1/2n$.
 
@@
 
-We conjecture, and numerically verify for $d\le200$, that measurements of spin projection along $2d-1$ axes suffice to reconstruct an arbitrary state of a $d$-level spin qudit (see Figure [\red{REF}]).
-Even so, $2d-1$ measurement axes may not be the ``optimal'' number to use.
+We conjecture, and numerically verified for $d\le200$, that measurements of spin projection along $2d-1$ axes suffice to reconstruct an arbitrary state of a $d$-level spin qudit.
+Even so, $2d-1$ measurement axes may not be the optimal number to use.
 Increasing the number of measurement axes generally decreases the error scale $\epsilon_V^2$, but at the cost of having to estimate more observables.
-At a fixed total number of measurements, $N=n\abs{V}$, increasing the number of measurement axes $\abs{V}$ reduces the number of measurements $n$ along each axis $\v v\in V$.
+At a fixed total number of measurements, increasing the number of measurement axes $\abs{V}$ reduces the number of measurements $n$ devoted to each axis $\v v\in V$.
 This trade-off begs the question: how many measurement axes should be chosen to perform spin qudit tomography?
 
-The reconstruction error bound in Eq.~\eqref{eq:bound} nominally provides a straightforward answer: at a fixed total number of measurements $N=n\abs{V}$, the number of measurement axes should be chosen to minimize $\epsilon_V^2/n\propto\epsilon_V^2\abs{V}$.
+The reconstruction error bound in Eq.~\eqref{eq:bound} nominally provides a straightforward answer: at a fixed total number of measurements, $N=n\abs{V}$, the number of measurement axes should be chosen to minimize $\epsilon_V^2/n\propto\epsilon_V^2\abs{V}$.
 

### 2020-11-09 sun_phases/scripts/compute_error_scales.py

@@ -150,4 +150,8 @@ for dim in range(min_dim, max_dim+1):
     print(mean_time, min_error_scale)
     sys.stdout.flush()
 
-print("total runtime:", time.time() - genesis)
+runtime = time.time() - genesis
+runtime_text = f"total runtime: {runtime}"
+print(runtime_text)
+with open(data_file, "a") as file:
+    file.write(f"# {runtime_text}\n")

### 2020-11-09 sun_phases/scripts/plot_bound_runtimes.py

@@ -1,0 +1,51 @@
+#!/usr/bin/env python3
+
+import os, sys, glob
+import numpy as np
+import matplotlib.pyplot as plt
+
+serial = "serial" in sys.argv
+
+data_dir = "../data/error_scale/"
+fig_dir = "../figures/error_scale/"
+
+if not os.path.isdir(fig_dir):
+    os.makedirs(fig_dir)
+
+params = { "font.size" : 10,
+           "text.usetex" : True }
+plt.rcParams.update(params)
+
+figsize = (3,2)
+fit_dim_min = 100
+
+##################################################
+
+files = glob.glob(data_dir + "times_*" + ("serial" if serial else "") + ".txt")
+
+def max_dim(file):
+    pieces = file.replace(".txt","").split("/")[-1].split("_")
+    return pieces[-1].split("-")[-1]
+files = sorted(files, key = max_dim)
+
+data = np.vstack([ np.loadtxt(file) for file in files ])
+sizes = data[:,0]
+times = data[:,1]
+
+plt.figure(figsize = figsize)
+plt.loglog(sizes, times, "k.", label = "data")
+
+size_lims = plt.gca().get_xlim()
+time_lims = plt.gca().get_ylim()
+
+fit_idx = sizes >= fit_dim_min
+fit = np.polyfit(np.log(sizes[fit_idx]), np.log(times[fit_idx]), deg = 1)
+plt.loglog(size_lims, np.exp(fit[0]*np.log(size_lims) + fit[1]), "k--", label = "fit")
+plt.gca().set_xlim(size_lims)
+plt.gca().set_ylim(time_lims)
+
+plt.xlabel(r"qudit dimension $d$")
+plt.ylabel(r"seconds to compute $\mathcal{S}_V$")
+plt.legend(loc = "best")
+plt.tight_layout()
+plt.savefig(fig_dir + "error_times.pdf")

### 2020-11-09 sun_phases/spin_qudit_tomo.tex
add some minor explanations

@@
 \end{align}
-where for any $d\times d$ matrix $X=\sum_{\mu,\nu} X_{\mu\nu} \op{\mu}{\nu}$ we define the $d^2$-component vector $\oket{X} \equiv \sum_{\mu,\nu} X_{\mu\nu} \ket{\mu\nu}$; $\obra{X}$ is the conjugate transpose of $\oket{X}$; and $\delta_{kk'}\equiv 1$ if $k=k'$ and $0$ otherwise.
+where for any $d\times d$ matrix $X=\sum_{\mu,\nu} X_{\mu\nu} \op{\mu}{\nu}$ we define the $d^2$-component vector $\oket{X} \equiv \sum_{\mu,\nu} X_{\mu\nu} \ket{\mu\nu}$; $\obra{X}$ is the conjugate transpose of $\oket{X}$, such that $\obk{X|Y}=\tr\p{X^\dag Y}$; and $\delta_{kk'}\equiv 1$ if $k=k'$ and $0$ otherwise.
 As a consequence, we can expand any density operator $\rho$ in the transition operator basis as
@@
 For any fixed axis $\v v$, the sets $\set{\Pi_{\v v\mu}}$ and $\set{T_{\v v\ell,0}}$ are both orthonormal bases for the space of operators that are diagonal in the basis $\set{\ket{\mu_{\v v}}}$.
-By a simple change of operator basis, measuring the projectors $\set{\Pi_{\v v\mu}}$ is therefore equivalent to measuring the transition operators $\set{T_{\v v\ell,0}}$, and provides data on the expectation values $\bk{T_{\v \ell,0}}_\rho$, where $\bk{X}_\rho\equiv\obk{\rho|X}$.
+By a simple change of operator basis, measuring the projectors $\set{\Pi_{\v v\mu}}$ is therefore equivalent to measuring the transition operators $\set{T_{\v v\ell,0}}$, and provides data on the expectation values $\bk{T_{\v v\ell,0}}_\rho$, where $\bk{X}_\rho\equiv\obk{\rho|X}$.
 
@@
 Unfortunately, we have no general proof of this conjecture, which requires a deeper analysis of fixed-degree signal recovery on the sphere.
-Nonetheless, we have numerically verified that $2d-1$ measurement axes suffice for all $d\le200$ (see Section \ref{sec:bound}).
+Nonetheless, in the following section we numerically verify that $2d-1$ measurement axes suffice for all $d\le200$ [\red{todo: possibly refer to Section \ref{sec:axes} instead}].
 
@@
   &&
-  t_{\ell\mu} \equiv \bk{\mu|T_{\ell,0}|\mu},
+  t_{\ell\mu} \equiv \bk{\mu|T_{\ell,0}|\mu}
+  = \sqrt{\f{2\ell+1}{d}} \bk{s\mu;\ell,0|s\mu},
 \end{align}

### 2020-11-09 sun_phases/spin_qudit_tomo.tex
change some defaults in error bound runtime scripts

@@
 Computing $\S_V^2$ to certify the bound in Eq.~\eqref{eq:bound_eps} requires constructing the measurement matrix $M_V$ and finding its singular values.
-This complexity of this task can be greatly reduced by the fact that the degree $\ell$ of the transition operator $T_{\ell,0}$ is preserved under rotations, which implies that the unitary
+The complexity of this task can be greatly reduced by the fact that the degree $\ell$ of the transition operator $T_{\ell,0}$ is preserved under rotations, which implies that the unitary
 \begin{align}

### 2020-11-09 sun_phases/scripts/compute_error_scales.py

@@ -12,10 +12,10 @@ except NotImplementedError:
 
 np.set_printoptions(linewidth = 200)
 
-parallelize = True
-if "serial" in sys.argv:
-    parallelize = False
-    sys.argv.remove("serial")
+parallelize = False
+if "par" in sys.argv:
+    parallelize = True
+    sys.argv.remove("par")
 
 min_dim = int(sys.argv[1])
 try: max_dim = int(sys.argv[2])

### 2020-11-09 sun_phases/scripts/plot_bound_runtimes.py

@@ -1,10 +1,8 @@
 #!/usr/bin/env python3
 
-import os, sys, glob
+import os, glob
 import numpy as np
 import matplotlib.pyplot as plt
-
-serial = "serial" in sys.argv
 
 data_dir = "../data/error_scale/"
 fig_dir = "../figures/error_scale/"
@@ -21,7 +19,7 @@ fit_dim_min = 100
 
 ##################################################
 
-files = glob.glob(data_dir + "times_*" + ("serial" if serial else "") + ".txt")
+files = glob.glob(data_dir + "times_*_serial.txt")
 
 def max_dim(file):
     pieces = file.replace(".txt","").split("/")[-1].split("_")

### 2020-11-09 sun_phases/spin_qudit_tomo.tex
minor cleaning up of some text

@@
 We then map this quantum tomography problem onto the well-studied classical problem of signal recovery on the sphere in Section \ref{sec:signal_recovery}, thereby importing a host of existing literature and mathematical machinery that has been developed for this task \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
-In particular, this mapping allows us to bound the minimum number of measurement axes necessary to perform full tomography on a $d$-level spin qudit, $r_d$, by $2d-1\le r_d\le d^2$.
+In particular, this mapping allows us to bound the minimum number of measurement axes necessary to perform full tomography on a $d$-level spin qudit, $r_d^{\t{min}}$, by $2d-1\le r_d^{\t{min}}\le d^2$.
 
@@
 where for any $d\times d$ matrix $X=\sum_{\mu,\nu} X_{\mu\nu} \op{\mu}{\nu}$ we define the $d^2$-component vector $\oket{X} \equiv \sum_{\mu,\nu} X_{\mu\nu} \ket{\mu\nu}$; $\obra{X}$ is the conjugate transpose of $\oket{X}$, such that $\obk{X|Y}=\tr\p{X^\dag Y}$; and $\delta_{kk'}\equiv 1$ if $k=k'$ and $0$ otherwise.
-As a consequence, we can expand any density operator $\rho$ in the transition operator basis as
+These properties of the transition operators allow us to expand any density operator $\rho$ in the transition operator basis as
 \begin{align}
-  \rho = \sum_{\ell=0}^{2s} \sum_{m=-\ell}^\ell \rho_{\ell m} T_{\ell m},
+  \rho = \sum_{\ell=0}^{d-1} \sum_{m=-\ell}^\ell \rho_{\ell m} T_{\ell m},
   &&
@@
 For a $d$-level spin, these quasi-probability distributions are band-limited at degree $d$.}.
-Moreover, for any fixed degree $\ell$, finding solutions to the reduced recovery problem in Eq.~\eqref{eq:tomo_recovery} is equivalent to the recovery of a degree-$\ell$ function $f = \sum_m f_m Y_{\ell m}$, which is possible with $\abs{V}=2\ell+1$ samples \cite{freeden2008spherical}.
+Moreover, for any fixed degree $\ell$, finding solutions to the reduced recovery problem in Eq.~\eqref{eq:tomo_recovery} is equivalent to the recovery of a degree-$\ell$ function $f = \sum_m f_m Y_{\ell m}$, which provably is possible with $\abs{V}=2\ell+1$ samples \cite{freeden2008spherical}.
 In the case of spin qudit tomography, the degree $\ell$ takes a maximal value of $\ell_{\t{max}}\equiv d-1$, so state recovery requires at least as many measurement axes as there are transition operators with degree $\ell_{\t{max}}$, namely $2\ell_{\t{max}}+1=2d-1$.
@@
 After choosing a set of measurement axes $V$, experimentally characterize the expectation values $\bk{\Pi_{\v v\mu}}_\rho$ by making measuring spin projection along every axis $\v v\in V$.
-These spin projections determine the expectation values $\bk{T_{\v v\ell,0}}_\rho$, whose estimates can be used to construct the state $\tilde\rho$ defined in Eq.~\eqref{eq:reconstructed_state}.
-If spin projection was measured $n$ times along every axis, then the reconstructed state $\tilde\rho$ approximates the true state $\rho$ to within a mean squared (Euclidean) distance $\epsilon_V^2/n$.
+These spin projections determine the expectation values $\bk{T_{\v v\ell,0}}_\rho$, whose estimates can be used to construct the state $\tilde\rho_V$ defined in Eq.~\eqref{eq:reconstructed_state}.
+If spin projection was measured $n$ times along every axis, then the reconstructed state $\tilde\rho_V$ approximates the true state $\rho$ to within a mean squared (Euclidean) distance $\epsilon_V^2/n$.
 Furthermore, if $\tilde\rho_V$ has negative eigenvalues, its distance from $\rho$ can be further reduced with maximum-likelihood corrections \cite{smolin2012efficient}.
@@
 Such a procedure is equivalent to making $N$ local measurements of the $N$-fold product state $\rho^{\otimes N}$.
-For convenience, we index the tensor factors of $\rho^{\otimes N}$ by the integers $\p{i,j}$, with $i\in\set{1,2,\cdots,\abs{V}}$ specifying a measurement axis $\v v_i\in V$ and $j\in\set{1,2,\cdots,n}$ specifying the copy of $\rho$ prepared for the $j$-th measurement spin projection along a particular axis.
-We then define the projectors $\Pi_{i\mu}\equiv\op{\mu_{\v v_i}}$ onto states $\ket{\mu_{\v v_i}}$ with definite spin projection $\mu$ onto axis $\v v_i\in V$, and define $\Pi_{i\mu}^j$ to be an $N$-qudit operator with $\Pi_{i\mu}$ on the $\p{i,j}$-th tensor factor and the identity elsewhere.
+For convenience, we index the tensor factors of $\rho^{\otimes N}$ by the integers $\p{i,j}$, with $i\in\set{1,2,\cdots,\abs{V}}$ specifying a measurement axis $\v v_i\in V$, and $j\in\set{1,2,\cdots,n}$ specifying the copy of $\rho$ prepared for the $j$-th measurement spin projection along a particular axis.
+We then define the projectors $\Pi_{i\mu}\equiv\op{\mu_{\v v_i}}$ onto states $\ket{\mu_{\v v_i}}$ with definite spin projection $\mu$ along axis $\v v_i\in V$, and define $\Pi_{i\mu}^j$ to be an $N$-qudit operator with $\Pi_{i\mu}$ on the $\p{i,j}$-th tensor factor and the identity elsewhere.
 We denote the experimental outcome of measuring $\Pi_{i\mu}$ in the $\p{i,j}$-th copy of $\rho$ by $\tilde \Pi_{i\mu}^j\in\set{0,1}$.
 In other words, $\tilde \Pi_{i\mu}^j$ is the ``single-shot estimate'' of $\Pi_{i\mu}$, with $\tilde \Pi_{i\mu}^j=1$ if outcome $\mu$ was observed on the $\p{i,j}$-th experimental trial, and $\tilde \Pi_{i\mu}^j=0$ otherwise.
-An empirical estimate of the expectation value $\bk{\Pi_{i\mu}}_\rho$ is then provided by the fraction of times that outcome $\mu$ was observed when measuring spin projection along axis $\v v_i$, that is
+An empirical estimate of the expectation value $\bk{\Pi_{i\mu}}_\rho$ is provided by the fraction of times that outcome $\mu$ was observed when measuring spin projection along axis $\v v_i$, that is
 \begin{align}
@@
 
-Due to finite sampling error (i.e.~shot noise), an empirical estimate $\tilde\O$ of the observable $\O$ will have statistical error with mean zero:
+Finite sampling error (i.e.~shot noise) generally induces statistical error $\epsilon_\O$ into the empirical estimate $\tilde\O$ of an observable $\O$:
 \begin{align}
@@
 If $\p{i,j}\ne\p{i',j'}$, then $\Pi_{i\mu}^j$ and $\Pi_{i'\mu'}^{j'}$ address different tensor factors of the product state $\rho^{\otimes N}$, so the expectation value of their product factorizes due to the identity $\tr\sp{\p{A\otimes B}\p{A'\otimes B'}} = \tr\p{AA'}\times\tr\p{BB'}$.
+This factorization can also be seen as a consequence of the fact that if $\p{i,j}\ne\p{i',j'}$, then $\Pi_{i\mu}^j$ and $\Pi_{i'\mu'}^{j'}$ are ``spatially separated'', which means that their expectation values with respect to the product state $\rho^{\otimes N}$ cannot have quantum correlations.
 The terms in Eq.~\eqref{eq:proj_cov_start} with $\p{i,j}\ne\p{i',j'}$ therefore vanish, so
@@
 and likewise $\tilde T_{i\ell} = \sum_\mu t_{\ell\mu} \tilde\Pi_{i\mu}$.
-The covariances between statistical errors in the transition operator basis are then
+The covariance between errors in the transition operator basis is then
 \begin{align}
@@
   = \delta_{ii'} \times \f1n \sum_{\mu,\mu'} t_{\ell\mu} t_{\ell'\mu'}
-  \sp{\delta_{\mu\mu'} p_\mu^i - p_\mu^i p_{\mu'}^i}.
+  \p{\delta_{\mu\mu'} p_\mu^i - p_\mu^i p_{\mu'}^i}.
 \end{align}
@@
 
-Though the result in Eq.~\eqref{eq:variance_bound} is sufficient to proceed with our derivation of an improved tomographic reconstruction error bound, we first take some time to better understand the bound on $\bbk{\epsilon_{i\ell}^2}$ analytically.
+Though the result in Eq.~\eqref{eq:variance_bound} is sufficient to proceed with our derivation of an improved tomographic reconstruction error bound, we first take some time to build a better analytical understanding of the bound on $\bbk{\epsilon_{i\ell}^2}$.
 To this end, we first note that the Clebsch-Gordan coefficients $\tau_{\ell\mu}$ are at most $1$ in magnitude, which implies that $\gamma_\ell^2\le 1$.
@@
 \begin{align}
-  \sigma_p^2\p{t_\ell} \le \sum_\mu t_{\ell\mu}^2 p_\mu
+  \sigma_p^2\p{t_\ell} \le \sum_\mu p_\mu t_{\ell\mu}^2
   \le \sum_\mu t_{\ell\mu}^2 = 1,
@@
 \begin{align}
-  \sigma_p^2\p{t_\ell} \le \sum_\mu t_{\ell\mu}^2 p_\mu
+  \sigma_p^2\p{t_\ell} \le \sum_\mu p_\mu t_{\ell\mu}^2
   \le t_{\ell\mu_{\t{max}}}^2
@@
   \bbk{\epsilon_{i\ell}^2}
-  < \f{2\ell+1}{nd} \times \gamma_\ell^2
+  \le \f{2\ell+1}{nd} \times \gamma_\ell^2
   \times \p{1-\delta_{\ell,0}},

### 2020-11-09 sun_phases/spin_qudit_tomo.tex
added comment about state-dependent error bound

@@
 
-[\red{todo: add a state-dependent error bound}]
+As a final comment, we note that the statistical error bound in Eq.~\eqref{eq:bound} is a ``worst case'' bound that is agnostic to the state $\rho$.
+The actual statistical error for any particular state may be considerably smaller.
+Written out in full (and derived in Appendix \ref{sec:error}),
+\begin{align}
+  \Bbk{\norm{\tilde\rho_V-\rho}^2}
+  = \f1n \sum_{i,\ell} \sigma_\rho^2\p{T_{\v v_i\ell,0}}
+  \sum_m \abs{q_{\ell mi}^V}^2,
+  \label{eq:error}
+\end{align}
+where $i$ indexes an axis $\v v_i\in V$; $\sigma_\rho^2\p{X} \equiv \bk{X^2}_\rho - \bk{X}_\rho^2$ is the variance of $X$; and $\v q_{\ell m}^V$ is a right singular vector of $M_{V\ell}$ corresponding to the singular value $M^V_{\ell m}$, with normalization $\abs{\v q_{\ell m}^V}=1/M^V_{\ell m}$.
+While the error in Eq.~\eqref{eq:error} cannot be known exactly without knowing $\rho$, this error can be estimated by computing the right-hand side of Eq.~\eqref{eq:error} with the replacement $\rho\to\tilde\rho_V$.
 
@@
   = \delta_{ii'} \times \f1n \sum_{\mu,\mu'} t_{\ell\mu} t_{\ell'\mu'}
-  \p{\delta_{\mu\mu'} p_\mu^i - p_\mu^i p_{\mu'}^i}.
+  \p{\delta_{\mu\mu'} p_\mu^i - p_\mu^i p_{\mu'}^i}
+  = \delta_{ii'} \times \f1n \sp{\bk{T_{i\ell} T_{i\ell'}}_\rho
+    - \bk{T_{i\ell}}_\rho \bk{T_{i\ell'}}_\rho}.
 \end{align}
@@
 
+As a final comment, we note that the statistical error bound in \eqref{eq:bound_full} is a ``worst-case'' bound that is agnostic to the state $\rho$.
+The actual statistical error for any particular state may be considerably smaller, and can be obtained by substituting the exact expression for $\bbk{\epsilon_{i\ell}^2}$:
+\begin{align}
+  \Bbk{\norm{\tilde\rho_V-\rho}^2}
+  = \sum_{i,\ell,m} \abs{q^V_{\ell mi}}^2 \bbk{\epsilon_{i\ell}^2}
+  = \f1n \sum_{i,\ell} \sigma_\rho^2\p{T_{i\ell}}
+  \sum_m \abs{q^V_{\ell mi}}^2,
+  \label{eq:error_full}
+\end{align}
+where $\sigma_\rho^2\p{X}\equiv\bk{X^2}_\rho-\bk{X}_\rho^2$ is the variance of $X$.
+While the error in Eq.~\eqref{eq:error_full} cannot be known exactly without knowing $\rho$, this error can be estimated by computing the right-hand side of Eq.~\eqref{eq:error_full} with the replacement $\rho\to\tilde\rho_V$.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2020-11-10 sun_phases/spin_qudit_tomo.tex
simplified exact expression for reconstruction error

@@
 \renewcommand{\S}{\mathcal{S}}
+\newcommand{\T}{\mathcal{T}}
 
@@
 \def\oop#1#2{\oket{#1}\!\obra{#2}}
+
+\DeclareMathOperator{\cov}{cov}
+\DeclareMathOperator{\diag}{diag}
 
@@
 The actual statistical error for any particular state may be considerably smaller.
-Written out in full (and derived in Appendix \ref{sec:error}),
+To write out the error in full, we identify the normalized right singular vectors $\v x^V_{\ell m}\equiv (x^V_{\ell m,1}, x^V_{\ell m,2}, \cdots)$ and corresponding singular values $M^V_{\ell m}$ of $M_{V\ell}$, and define
+\begin{align}
+  C_{V\ell} \equiv \sum_i C_{V\ell i} \op{i},
+  &&
+  C_{V\ell i} \equiv \f1{\p{M^V_{\ell m}}^2} \sum_m \abs{x_{\ell m i}}^2.
+\end{align}
+We can then compactly express (see Appendix \ref{sec:error})
 \begin{align}
   \Bbk{\norm{\tilde\rho_V-\rho}^2}
-  = \f1n \sum_{i,\ell} \sigma_\rho^2\p{T_{\v v_i\ell,0}}
-  \sum_m \abs{q_{\ell mi}^V}^2,
+  = \f1n \sum_\ell \obk{\xi_{V\ell}|\T_{\rho\ell}},
   \label{eq:error}
 \end{align}
-where $i$ indexes an axis $\v v_i\in V$; $\sigma_\rho^2\p{X} \equiv \bk{X^2}_\rho - \bk{X}_\rho^2$ is the variance of $X$; and $\v q_{\ell m}^V$ is a right singular vector of $M_{V\ell}$ corresponding to the singular value $M^V_{\ell m}$, with normalization $\abs{\v q_{\ell m}^V}=1/M^V_{\ell m}$.
+where
+\begin{align}
+  \xi_{V\ell} &\equiv M_{V\ell} C_{V\ell} M_{V\ell}^\dag, \\
+  \T_{\rho\ell} &\equiv \sum_{m,n}
+  \cov_\rho\p{T_{\ell m}^\dag, T_{\ell n}} \op{m}{n},
+\end{align}
+with $\cov_\rho\p{X,Y} \equiv \bk{XY}_\rho - \bk{X}_\rho \bk{Y}_\rho$ the covariance between $X$ and $Y$.
 While the error in Eq.~\eqref{eq:error} cannot be known exactly without knowing $\rho$, this error can be estimated by computing the right-hand side of Eq.~\eqref{eq:error} with the replacement $\rho\to\tilde\rho_V$.
@@
   = \sum_{i,\ell,m} \abs{q^V_{\ell mi}}^2 \bbk{\epsilon_{i\ell}^2}
-  = \f1n \sum_{i,\ell} \sigma_\rho^2\p{T_{i\ell}}
-  \sum_m \abs{q^V_{\ell mi}}^2,
+  = \f1n \sum_{i,\ell} C^V_{\ell i} \sigma_\rho^2\p{T_{i\ell}},
+  &&
+  C^V_{\ell i} \equiv \sum_m \abs{q_{\ell mi}}^2,
+  \label{eq:error_full_start}
+\end{align}
+where $\sigma_\rho^2\p{X}\equiv\bk{X^2}_\rho - \bk{X}_\rho^2$ is the variance of $X$.
+To simplify this error, we use the fact that all $T_{i\ell}=T_{i\ell}^\dag$ to expand
+\begin{align}
+  \sigma_\rho^2\p{T_{i\ell}}
+  = \sigma_\rho^2\p{\sum_m D^\ell_{m,0}\p{\v v_i} T_{\ell m}}
+  = \sum_{m,n} D^\ell_{m,0}\p{\v v_i}^* D^\ell_{n,0}\p{\v v_i}
+  \cov_\rho\p{T_{\ell m}^\dag, T_{\ell n}},
+\end{align}
+where
+\begin{align}
+  \cov_\rho\p{X,Y} \equiv \bk{X Y}_\rho - \bk{X}_\rho \bk{Y}_\rho
+\end{align}
+is the covariance between $X$ and $Y$.
+This expansion allows us to express the error in Eq.~\eqref{eq:error_full_start} more compactly as
+\begin{align}
+  \Bbk{\norm{\tilde\rho_V-\rho}^2}
+  = \f1n \sum_\ell \p{\xi_{V\ell}|\T_{\rho\ell}},
   \label{eq:error_full}
 \end{align}
-where $\sigma_\rho^2\p{X}\equiv\bk{X^2}_\rho-\bk{X}_\rho^2$ is the variance of $X$.
+where
+\begin{align}
+  \xi_{V\ell} \equiv M_{V\ell} C_{V\ell} M_{V\ell}^\dag,
+  &&
+  C_{V\ell}
+  \equiv \diag\p{C^V_{\ell,1},C^V_{\ell,2},\cdots,C^V_{\ell,\abs{V}}}
+\end{align}
+and
+\begin{align}
+  \T_{\rho\ell}
+  \equiv \sum_{m,n} \cov_\rho\p{T_{\ell m}^\dag, T_{\ell n}} \op{m}{n}
+  = \sum_{m,n} \sp{\delta_{mn} \bk{T_{\ell m}^\dag T_{\ell m}}_\rho
+    - \bk{T_{\ell m}^\dag}_\rho \bk{T_{\ell n}}_\rho} \op{m}{n}.
+\end{align}
 While the error in Eq.~\eqref{eq:error_full} cannot be known exactly without knowing $\rho$, this error can be estimated by computing the right-hand side of Eq.~\eqref{eq:error_full} with the replacement $\rho\to\tilde\rho_V$.

### 2020-11-10 sun_phases/spin_qudit_tomo.tex
renamed some equations and fixed an error in Eq. B32

@@
   < \f1{nd} \sum_{\ell>0} \p{2\ell+1} \gamma_\ell^2 \S_{V\ell}^2.
-  \label{eq:bound_full}
+  \label{eq:bound_SM}
 \end{align}
-We note that whereas the bound on $\bbk{\epsilon_{i\ell}^2}$ in Eq.~\eqref{eq:variance_bound} is tight, the bounds in Eq.~\eqref{eq:bound_single} and Eq.~\eqref{eq:bound_full} are not: the upper bound on $\bbk{\epsilon_{i\ell}^2}$ cannot be achieved simultaneously for all axes $\v v_i$ or degrees $\ell$.
-In principle, there is therefore room for improvement on the bound in Eq.~\eqref{eq:bound_full}.
+We note that whereas the bound on $\bbk{\epsilon_{i\ell}^2}$ in Eq.~\eqref{eq:variance_bound} is tight, the bounds in Eq.~\eqref{eq:bound_single} and Eq.~\eqref{eq:bound_SM} are not: the upper bound on $\bbk{\epsilon_{i\ell}^2}$ cannot be achieved simultaneously for all axes $\v v_i$ or degrees $\ell$.
+In principle, there is therefore room for improvement on the bound in Eq.~\eqref{eq:bound_SM}.
 
-As a final comment, we note that the statistical error bound in \eqref{eq:bound_full} is a ``worst-case'' bound that is agnostic to the state $\rho$.
+As a final comment, we note that the statistical error bound in \eqref{eq:bound_SM} is a ``worst-case'' bound that is agnostic to the state $\rho$.
 The actual statistical error for any particular state may be considerably smaller, and can be obtained by substituting the exact expression for $\bbk{\epsilon_{i\ell}^2}$:
@@
   C^V_{\ell i} \equiv \sum_m \abs{q_{\ell mi}}^2,
-  \label{eq:error_full_start}
+  \label{eq:error_SM_start}
 \end{align}
@@
 is the covariance between $X$ and $Y$.
-This expansion allows us to express the error in Eq.~\eqref{eq:error_full_start} more compactly as
+This expansion allows us to express the error in Eq.~\eqref{eq:error_SM_start} more compactly as
 \begin{align}
@@
   = \f1n \sum_\ell \p{\xi_{V\ell}|\T_{\rho\ell}},
-  \label{eq:error_full}
+  \label{eq:error_SM}
 \end{align}
@@
   C_{V\ell}
-  \equiv \diag\p{C^V_{\ell,1},C^V_{\ell,2},\cdots,C^V_{\ell,\abs{V}}}
+  \equiv \diag\p{C^V_{\ell,1},C^V_{\ell,2},\cdots,C^V_{\ell,\abs{V}}},
+  &&
+  \T_{\rho\ell}
+  \equiv \sum_{m,n} \cov_\rho\p{T_{\ell m}^\dag, T_{\ell n}} \op{m}{n}.
 \end{align}
-and
-\begin{align}
-  \T_{\rho\ell}
-  \equiv \sum_{m,n} \cov_\rho\p{T_{\ell m}^\dag, T_{\ell n}} \op{m}{n}
-  = \sum_{m,n} \sp{\delta_{mn} \bk{T_{\ell m}^\dag T_{\ell m}}_\rho
-    - \bk{T_{\ell m}^\dag}_\rho \bk{T_{\ell n}}_\rho} \op{m}{n}.
-\end{align}
-While the error in Eq.~\eqref{eq:error_full} cannot be known exactly without knowing $\rho$, this error can be estimated by computing the right-hand side of Eq.~\eqref{eq:error_full} with the replacement $\rho\to\tilde\rho_V$.
+While the error in Eq.~\eqref{eq:error_SM} cannot be known exactly without knowing $\rho$, this error can be estimated by computing the right-hand side of Eq.~\eqref{eq:error_SM} with the replacement $\rho\to\tilde\rho_V$.
 

### 2020-11-10 sun_phases/spin_qudit_tomo.tex
fixed error in Eq. 31
added comment about tightness to the main text

@@
 If $d$ even or $\ell$ is odd, then $\gamma_\ell=\max_\mu\tau_{\ell\mu}$.
-For comparison with the previous bound in Eq.~\eqref{eq:bound_eps}, we note that the error scale $\epsilon_V^2<\S_V^2/2$, so the previous bound still holds with the replacement $\epsilon^2\to1/2n$.
+For comparison with the previous bound in Eq.~\eqref{eq:bound_eps}, the error scale $\epsilon_V^2<\S_V^2/2$, so the previous bound still holds with the replacement $\epsilon^2\to1/2n$.
+We note that the bound in Eq.~\eqref{eq:bound} is not tight, as it is acquired by bounding the statistical variance $\epsilon_{\v v\ell}^2$ in the empirical estimate of $\bk{T_{\v v\ell,0}}_\rho$ as $\bbk{\epsilon_{\v v\ell}^2}\le\frac{2\ell+1}{nd}\times\gamma_\ell^2$.
+Though the individual bounds for each axis $\v v$ and degree $\ell$ are tight, these bounds cannot be achieved simultaneously.
+There is therefore still room for improvement on the bound in Eq.~\eqref{eq:bound}.
 
-As a final comment, we note that the statistical error bound in Eq.~\eqref{eq:bound} is a ``worst case'' bound that is agnostic to the state $\rho$.
-The actual statistical error for any particular state may be considerably smaller.
-To write out the error in full, we identify the normalized right singular vectors $\v x^V_{\ell m}\equiv (x^V_{\ell m,1}, x^V_{\ell m,2}, \cdots)$ and corresponding singular values $M^V_{\ell m}$ of $M_{V\ell}$, and define
+As a final comment, we note that Eq.~\eqref{eq:bound} provides a pessimistic upper bound on statistical error, which can be calculated without prior knowledge of the true qudit state $\rho$.
+The actual error in the reconstruction $\tilde\rho_V$ of a particular state $\rho$ may be considerably smaller.
+To write out this error in full, we identify the normalized right singular vectors $\v x^V_{\ell m}\equiv (x^V_{\ell m,1}, x^V_{\ell m,2}, \cdots)$ and corresponding singular values $M^V_{\ell m}$ of $M_{V\ell}$, and define
 \begin{align}
-  C_{V\ell} \equiv \sum_i C_{V\ell i} \op{i},
+  C_{V\ell i} \equiv \sum_m \abs{\f{x_{\ell m i}}{M^V_{\ell m}}}^2,
   &&
-  C_{V\ell i} \equiv \f1{\p{M^V_{\ell m}}^2} \sum_m \abs{x_{\ell m i}}^2.
+  C_{V\ell} \equiv \sum_i C_{V\ell i} \op{\v v_i}.
 \end{align}
@@
 \end{align}
-with $\cov_\rho\p{X,Y} \equiv \bk{XY}_\rho - \bk{X}_\rho \bk{Y}_\rho$ the covariance between $X$ and $Y$.
+with $\cov_\rho\p{X,Y} \equiv \bk{XY}_\rho - \bk{X}_\rho \bk{Y}_\rho$.
 While the error in Eq.~\eqref{eq:error} cannot be known exactly without knowing $\rho$, this error can be estimated by computing the right-hand side of Eq.~\eqref{eq:error} with the replacement $\rho\to\tilde\rho_V$.
@@
   \bbk{\epsilon_{i\ell}^2}
-  \le \f{2\ell+1}{nd} \times \gamma_\ell^2
+  < \f{2\ell+1}{nd} \times \gamma_\ell^2
   \times \p{1-\delta_{\ell,0}},

### 2020-11-10 sun_phases/spin_qudit_tomo.tex
fixed minor errors
improvements to the full expression of the statistical error in a reconstructed state

@@
 The actual error in the reconstruction $\tilde\rho_V$ of a particular state $\rho$ may be considerably smaller.
-To write out this error in full, we identify the normalized right singular vectors $\v x^V_{\ell m}\equiv (x^V_{\ell m,1}, x^V_{\ell m,2}, \cdots)$ and corresponding singular values $M^V_{\ell m}$ of $M_{V\ell}$, and define
+To write out this error in full (see Appendix \ref{sec:error} for a derivation), we identify the normalized right singular vectors $\v x^V_{\ell m}\equiv (x^V_{\ell m,1}, x^V_{\ell m,2}, \cdots)$ and corresponding singular values $M^V_{\ell m}$ of $M_{V\ell}$, and define
 \begin{align}
@@
 \end{align}
-We can then compactly express (see Appendix \ref{sec:error})
+We then construct the $\p{2\ell+1}\times\p{2\ell+1}$ matrix
+\begin{align}
+  \xi_{V\ell} \equiv M_{V\ell} C_{V\ell} M_{V\ell}^\dag,
+\end{align}
+whose (normalized) eigenvectors $\v\chi^V_{\ell m}$ determine the orthonormal operators
+\begin{align}
+  \Q_{V\ell m} \equiv \sum_n \chi^V_{\ell mn} T_{\ell n}.
+\end{align}
+In terms of the corresponding eigenvalues $\xi^V_{\ell m}$ of $\xi_{V\ell}$, we can then compactly express
 \begin{align}
   \Bbk{\norm{\tilde\rho_V-\rho}^2}
-  = \f1n \sum_\ell \obk{\xi_{V\ell}|\T_{\rho\ell}},
+  = \f1n \sum_{\ell,m} \xi^V_{\ell m}
+  \cov_\rho\p{\Q_{V\ell m}^\dag,\Q_{V\ell m}},
   \label{eq:error}
 \end{align}
-where
-\begin{align}
-  \xi_{V\ell} &\equiv M_{V\ell} C_{V\ell} M_{V\ell}^\dag, \\
-  \T_{\rho\ell} &\equiv \sum_{m,n}
-  \cov_\rho\p{T_{\ell m}^\dag, T_{\ell n}} \op{m}{n},
-\end{align}
-with $\cov_\rho\p{X,Y} \equiv \bk{XY}_\rho - \bk{X}_\rho \bk{Y}_\rho$.
+where $\cov_\rho\p{X,Y} \equiv \bk{XY}_\rho - \bk{X}_\rho \bk{Y}_\rho$.
 While the error in Eq.~\eqref{eq:error} cannot be known exactly without knowing $\rho$, this error can be estimated by computing the right-hand side of Eq.~\eqref{eq:error} with the replacement $\rho\to\tilde\rho_V$.
@@
   \bbk{\epsilon_{i\ell}^2}
-  < \f{2\ell+1}{nd} \times \gamma_\ell^2
+  < \f1{\p{M^V_{\ell m}}^2} \times \f{2\ell+1}{nd} \times \gamma_\ell^2
   \times \p{1-\delta_{\ell,0}},
@@
   = \f1n \sum_\ell \p{\xi_{V\ell}|\T_{\rho\ell}},
-  \label{eq:error_SM}
+  \label{eq:error_SM_mid}
 \end{align}
@@
   \equiv \sum_{m,n} \cov_\rho\p{T_{\ell m}^\dag, T_{\ell n}} \op{m}{n}.
+\end{align}
+The result in Eq.~\eqref{eq:error_SM_mid} can be further simplified by diagonalizing $\xi_{V\ell}$.
+Denoting the eigenvalues and corresponding eigenvectors of $\xi_{V\ell}$ respectively by $\xi^V_{\ell m}$ and $\v\chi^V_{\ell m}$, we finally arrive at the expression
+\begin{align}
+  \Bbk{\norm{\tilde\rho_V-\rho}^2}
+  = \f1n \sum_{\ell,m} \xi^V_{\ell m}
+  \cov_\rho\p{\Q_{V\ell m}^\dag, \Q_{V\ell m}},
+  &&
+  \Q_{V\ell m} \equiv \sum_n \chi^V_{\ell m n} T_{\ell n}.
+  \label{eq:error_SM}
 \end{align}

### 2020-11-11 sun_phases/spin_qudit_tomo.tex
simplified expression for exact reconstruction error

--- 2020-11-10 sun_phases/spin_qudit_tomo.tex
@@
 We note that the bound in Eq.~\eqref{eq:bound} is not tight, as it is acquired by bounding the statistical variance $\epsilon_{\v v\ell}^2$ in the empirical estimate of $\bk{T_{\v v\ell,0}}_\rho$ as $\bbk{\epsilon_{\v v\ell}^2}\le\frac{2\ell+1}{nd}\times\gamma_\ell^2$.
-Though the individual bounds for each axis $\v v$ and degree $\ell$ are tight, these bounds cannot be achieved simultaneously.
+Though the individual bounds for each axis $\v v$ and degree $\ell$
+are tight, these bounds cannot all be achieved simultaneously.
 There is therefore still room for improvement on the bound in Eq.~\eqref{eq:bound}.
@@
 The actual error in the reconstruction $\tilde\rho_V$ of a particular state $\rho$ may be considerably smaller.
-To write out this error in full, we identify the normalized right singular vectors $\v x^V_{\ell m}\equiv (x^V_{\ell m,1}, x^V_{\ell m,2}, \cdots)$ and corresponding singular values $M^V_{\ell m}$ of $M_{V\ell}$, and define
+To write out this error in full (see Appendix \ref{sec:error} for a derivation), we identify the singular value decomposition
 \begin{align}
-  C_{V\ell i} \equiv \sum_m \abs{\f{x_{\ell m i}}{M^V_{\ell m}}}^2,
-  &&
-  C_{V\ell} \equiv \sum_i C_{V\ell i} \op{\v v_i}.
+  M_{V\ell} = U_{V\ell} D_{V\ell} W_{V\ell}^\dag,
 \end{align}
-We can then compactly express (see Appendix \ref{sec:error})
+and define the $\p{2\ell+1}\times\p{2\ell+1}$ matrices
 \begin{align}
-  \Bbk{\norm{\tilde\rho_V-\rho}^2}
-  = \f1n \sum_\ell \obk{\xi_{V\ell}|\T_{\rho\ell}},
-  \label{eq:error}
-\end{align}
-where
-\begin{align}
-  \xi_{V\ell} &\equiv M_{V\ell} C_{V\ell} M_{V\ell}^\dag, \\
+  \xi_{V\ell} &\equiv M_{V\ell} \diag\p{W_{V\ell} D_{V\ell}^{-2} W_{V\ell}^\dag} M_{V\ell}^\dag, \\
   \T_{\rho\ell} &\equiv \sum_{m,n}
@@
 \end{align}
-with $\cov_\rho\p{X,Y} \equiv \bk{XY}_\rho - \bk{X}_\rho \bk{Y}_\rho$.
+where $\diag\p{X}$ sets all off-diagonal entries of $X$ to zero, and $\cov_\rho\p{X,Y} \equiv \bk{XY}_\rho - \bk{X}_\rho \bk{Y}_\rho$.
+We can then compactly express
+\begin{align}
+  \Bbk{\norm{\tilde\rho_V-\rho}^2}
+  = \f1n \sum_\ell \p{\xi_{V\ell}|\T_{\rho\ell}}.
+  \label{eq:error}
+\end{align}
 While the error in Eq.~\eqref{eq:error} cannot be known exactly without knowing $\rho$, this error can be estimated by computing the right-hand side of Eq.~\eqref{eq:error} with the replacement $\rho\to\tilde\rho_V$.
+
+[\red{I think the error in Eq.~\eqref{eq:error} can be maximized with no added cost to computational complexity.
+  If so, the second half of this section entire section (as well as Appendix \ref{sec:error}) should be rewritten to replace Eq.~\eqref{eq:bound} by the tighter bound obtained by maximizing Eq.~\eqref{eq:error}.}]
 
@@
   \p{\delta_{\mu\mu'} p_\mu^i - p_\mu^i p_{\mu'}^i}
-  = \delta_{ii'} \times \f1n \sp{\bk{T_{i\ell} T_{i\ell'}}_\rho
-    - \bk{T_{i\ell}}_\rho \bk{T_{i\ell'}}_\rho}.
+  = \delta_{ii'} \times \f1n \cov_\rho\p{T_{i\ell}, T_{i\ell'}},
 \end{align}
+where $\cov_\rho\p{X,Y} \equiv \bk{XY}_\rho - \bk{X}_\rho \bk{Y}_\rho$.
 In order to establish an upper bound on reconstruction error for an unknown quantum state, we will need an upper bound on the variance $\bbk{\epsilon_{i\ell}^2}$ in particular.
@@
 \end{align}
-is the variance of $X$ when sampled according to the probability distribution $p$.
+is the variance of the entries of $X$ when sampled according to the (classical) probability distribution $p$.
 The variance $\sigma_p^2\p{X}$ is maximal when $p$ has equal weight on the largest and smallest values of $X$, which implies that
@@
   = \sum_{i,\ell,m} \abs{q^V_{\ell mi}}^2 \bbk{\epsilon_{i\ell}^2}
-  = \f1n \sum_{i,\ell} C^V_{\ell i} \sigma_\rho^2\p{T_{i\ell}},
-  &&
-  C^V_{\ell i} \equiv \sum_m \abs{q_{\ell mi}}^2,
-  \label{eq:error_SM_start}
+  = \f1n \sum_{i,\ell} C^V_{\ell i} \cov_\rho\p{T_{i\ell},T_{i\ell}},
 \end{align}
-where $\sigma_\rho^2\p{X}\equiv\bk{X^2}_\rho - \bk{X}_\rho^2$ is the variance of $X$.
-To simplify this error, we use the fact that all $T_{i\ell}=T_{i\ell}^\dag$ to expand
+where
 \begin{align}
-  \sigma_\rho^2\p{T_{i\ell}}
-  = \sigma_\rho^2\p{\sum_m D^\ell_{m,0}\p{\v v_i} T_{\ell m}}
+  C^V_{\ell i} \equiv \sum_m \abs{q_{\ell mi}}^2
+  = \bk{\v v_i| W_{V\ell} D_{V\ell}^{-2} W_{V\ell}^\dag | \v v_i}
+\end{align}
+can be defined in terms of the singular value decomposition
+\begin{align}
+  M_{V\ell} = U_{V\ell} D_{V\ell} W_{V\ell}^\dag,
+\end{align}
+with a unitary matrix $U_{V\ell}$, diagonal matrix $D_{V\ell}$, and isometric matrix $W_{V\ell}$.
+Using the fact that all $T_{i\ell}=T_{i\ell}^\dag$, we can expand
+\begin{align}
+  \cov_\rho\p{T_{i\ell},T_{i\ell}}
+  = \cov_\rho\p{T_{i\ell}^\dag,T_{i\ell}}
   = \sum_{m,n} D^\ell_{m,0}\p{\v v_i}^* D^\ell_{n,0}\p{\v v_i}
@@
 \end{align}
-where
-\begin{align}
-  \cov_\rho\p{X,Y} \equiv \bk{X Y}_\rho - \bk{X}_\rho \bk{Y}_\rho
-\end{align}
-is the covariance between $X$ and $Y$.
-This expansion allows us to express the error in Eq.~\eqref{eq:error_SM_start} more compactly as
+which implies that
 \begin{align}
   \Bbk{\norm{\tilde\rho_V-\rho}^2}
-  = \f1n \sum_\ell \p{\xi_{V\ell}|\T_{\rho\ell}},
+  = \f1n \sum_{\ell,m,n,i}
+  D^\ell_{m,0}\p{\v v_i}^* D^\ell_{n,0}\p{\v v_i} C^V_{\ell i}
+  \cov_\rho\p{T_{\ell m}^\dag, T_{\ell n}}.
+\end{align}
+Altogether, we can define the $\p{2\ell+1}\times\p{2\ell+1}$ matrices
+\begin{align}
+  \xi_{V\ell} \equiv M_{V\ell}
+  \diag\p{W_{V\ell} D_{V\ell}^{-2} W_{V\ell}^\dag} M_{V\ell}^\dag
+  &&
+  \T_{\rho\ell} \equiv
+  \sum_{m,n} \cov_\rho\p{T_{\ell m}^\dag, T_{\ell n}} \op{m}{n},
+\end{align}
+where $\diag\p{X}$ sets all off-diagonal entries of $X$ to zero, and compactly express
+\begin{align}
+  \Bbk{\norm{\tilde\rho_V-\rho}^2}
+  = \f1n \sum_\ell \p{\xi_{V\ell}|\T_{\rho\ell}}.
   \label{eq:error_SM}
-\end{align}
-where
-\begin{align}
-  \xi_{V\ell} \equiv M_{V\ell} C_{V\ell} M_{V\ell}^\dag,
-  &&
-  C_{V\ell}
-  \equiv \diag\p{C^V_{\ell,1},C^V_{\ell,2},\cdots,C^V_{\ell,\abs{V}}},
-  &&
-  \T_{\rho\ell}
-  \equiv \sum_{m,n} \cov_\rho\p{T_{\ell m}^\dag, T_{\ell n}} \op{m}{n}.
 \end{align}

### 2020-11-11 sun_phases/spin_qudit_tomo.tex
simplify expressions for reconstruction error

@@
 \begin{align}
-  \S_V^2 \equiv \norm{M_V^{-1}}^2 = \sum_k\f1{\p{M^V_k}^2},
+  \S_V^2 \equiv \obk{M_V^{-1}|M_V^{-1}} = \sum_k\f1{\p{M^V_k}^2},
 \end{align}
@@
   &&
-  \S_{V\ell}^2 \equiv \norm{M_{V\ell}^{-1}}^2,
+  \S_{V\ell}^2 \equiv \obk{M_{V\ell}^{-1}|M_{V\ell}^{-1}},
 \end{align}
@@
 The actual error in the reconstruction $\tilde\rho_V$ of a particular state $\rho$ may be considerably smaller.
-To write out this error in full (see Appendix \ref{sec:error} for a derivation), we identify the singular value decomposition
+To write out this error in full (see Appendix \ref{sec:error} for a derivation), we define the $\p{2\ell+1}\times\p{2\ell+1}$ matrices
 \begin{align}
-  M_{V\ell} = U_{V\ell} D_{V\ell} W_{V\ell}^\dag,
-\end{align}
-and define the $\p{2\ell+1}\times\p{2\ell+1}$ matrices
-\begin{align}
-  \xi_{V\ell} &\equiv M_{V\ell} \diag\p{W_{V\ell} D_{V\ell}^{-2} W_{V\ell}^\dag} M_{V\ell}^\dag, \\
+  \xi_{V\ell} &\equiv M_{V\ell}
+  \diag\sp{M_{V\ell}^{-1} \p{M_{V\ell}^{-1}}^\dag} M_{V\ell}^\dag, \\
   \T_{\rho\ell} &\equiv \sum_{m,n}
@@
 \end{align}
-where $\diag\p{X}$ sets all off-diagonal entries of $X$ to zero, and $\cov_\rho\p{X,Y} \equiv \bk{XY}_\rho - \bk{X}_\rho \bk{Y}_\rho$.
-We can then compactly express
+where $\diag\sp{X}$ sets all off-diagonal entries of $X$ to zero, and $\cov_\rho\p{X,Y} \equiv \bk{XY}_\rho - \bk{X}_\rho \bk{Y}_\rho$.
+We discuss details of efficiently computing $\xi_{V\ell}$ in Appendix \ref{sec:compute}.
+These matrices allow us to compactly express
 \begin{align}
@@
 \end{align}
-The block diagonal structure of this matrix implies that, rather than an integer index $k\in\set{1,2,\cdots,d^2}$, we can identify the singular values and right singular vectors of $UM_V$ (and equivalently $M_V$) by a block index $\ell\in\set{0,1,\cdots,d-1}$ and a sub-index $m=\set{-\ell,-\ell+1,\cdots,\ell}$.
+The block diagonal structure of this matrix implies that, rather than an integer index $k\in\set{1,2,\cdots,d^2}$, we can identify the singular values and right singular vectors of $M_V$ by a block index $\ell\in\set{0,1,\cdots,d-1}$ and a sub-index $m=\set{-\ell,-\ell+1,\cdots,\ell}$.
 Making the replacement $k\to\p{\ell,m}$, we can thus repeat the derivations through Eq.~\eqref{eq:bound_eps} of Section \ref{sec:bound} without the assumption of Eq.~\eqref{eq:error_assumption}, finding that
@@
   = \sum_{i,\ell,m} \abs{q^V_{\ell mi}}^2 \bbk{\epsilon_{i\ell}^2}
-  = \f1n \sum_{i,\ell} C^V_{\ell i} \cov_\rho\p{T_{i\ell},T_{i\ell}},
+  = \f1n \sum_{\ell,i} C^V_{\ell i} \cov_\rho\p{T_{i\ell},T_{i\ell}},
+  &&
+  C^V_{\ell i} \equiv \sum_m \abs{q_{\ell mi}}^2.
+  \label{eq:error_SM_start}
 \end{align}
-where
+Identifying the singular value decomposition $M_{V\ell} = U_{V\ell} \Sigma_{V\ell} W_{V\ell}^\dag$, the vector $\tilde{\v q}_{\ell i} = (q_{\ell,-\ell,i},q_{\ell,-\ell+1,i},\cdots q_{\ell,\ell,i})$ is essentially the $i$-th row of $\Sigma_{V\ell}^{-1} W_{V\ell}^\dag$, which is to say that $\tilde{\v q}_{\ell i} = \Sigma_{V\ell}^{-1} W_{V\ell}^\dag \ket{\v v_i}$.
+We can therefore simplify
 \begin{align}
-  C^V_{\ell i} \equiv \sum_m \abs{q_{\ell mi}}^2
-  = \bk{\v v_i| W_{V\ell} D_{V\ell}^{-2} W_{V\ell}^\dag | \v v_i}
+  C^V_{\ell i}
+  = \abs{\tilde{\v q}_{\ell i}}^2
+  = \bk{\v v_i | \p{\Sigma_{V\ell}^{-1} W_{V\ell}^\dag}^\dag
+    \p{\Sigma_{V\ell}^{-1} W_{V\ell}^\dag} | \v v_i}
+  = \bk{\v v_i | W_{V\ell} \Sigma_{V\ell}^{-2} W_{V\ell}^\dag | \v v_i}
+  = \bk{\v v_i| M_{V\ell}^{-1} \p{M_{V\ell}^{-1}}^\dag | \v v_i}.
 \end{align}
-can be defined in terms of the singular value decomposition
-\begin{align}
-  M_{V\ell} = U_{V\ell} D_{V\ell} W_{V\ell}^\dag,
-\end{align}
-with a unitary matrix $U_{V\ell}$, diagonal matrix $D_{V\ell}$, and isometric matrix $W_{V\ell}$.
-Using the fact that all $T_{i\ell}=T_{i\ell}^\dag$, we can expand
+Using the fact that all $T_{i\ell}=T_{i\ell}^\dag$, we can also expand
 \begin{align}
@@
   \Bbk{\norm{\tilde\rho_V-\rho}^2}
-  = \f1n \sum_{\ell,m,n,i}
-  D^\ell_{m,0}\p{\v v_i}^* D^\ell_{n,0}\p{\v v_i} C^V_{\ell i}
+  = \f1n \sum_{i,\ell,m,n}
+  C^V_{\ell i} D^\ell_{m,0}\p{\v v_i}^* D^\ell_{n,0}\p{\v v_i}
   \cov_\rho\p{T_{\ell m}^\dag, T_{\ell n}}.
 \end{align}
-Altogether, we can define the $\p{2\ell+1}\times\p{2\ell+1}$ matrices
+Altogether, the result in Eq.~\eqref{eq:error_SM_start} can be expressed more compactly using the $\p{2\ell+1}\times\p{2\ell+1}$ matrices
 \begin{align}
-  \xi_{V\ell} \equiv M_{V\ell}
-  \diag\p{W_{V\ell} D_{V\ell}^{-2} W_{V\ell}^\dag} M_{V\ell}^\dag
+  \xi_{V\ell} \equiv
+  M_{V\ell} \diag\sp{M_{V\ell}^{-1} \p{M_{V\ell}^{-1}}^\dag} M_{V\ell}^\dag,
   &&
@@
 \end{align}
-where $\diag\p{X}$ sets all off-diagonal entries of $X$ to zero, and compactly express
+where $\diag\sp{X}$ sets all off-diagonal entries of $X$ to zero, in terms of which
 \begin{align}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Computing singular values of the measurement matrix}
+\section{Efficiently computing error bounds}
 \label{sec:compute}

### 2020-11-12 sun_phases/spin_qudit_tomo.tex
read through and simplified main text

@@
 For any fixed axis $\v v$, the sets $\set{\Pi_{\v v\mu}}$ and $\set{T_{\v v\ell,0}}$ are both orthonormal bases for the space of operators that are diagonal in the basis $\set{\ket{\mu_{\v v}}}$.
-By a simple change of operator basis, measuring the projectors $\set{\Pi_{\v v\mu}}$ is therefore equivalent to measuring the transition operators $\set{T_{\v v\ell,0}}$, and provides data on the expectation values $\bk{T_{\v v\ell,0}}_\rho$, where $\bk{X}_\rho\equiv\obk{\rho|X}$.
+By a simple change of operator basis, measuring the projectors $\set{\Pi_{\v v\mu}}$ is therefore equivalent to measuring the transition operators $\set{T_{\v v\ell,0}}$, and provides data on the expectation values $\bk{T_{\v v\ell,0}}_\rho$, where $\bk{X}_\rho \equiv \obk{\rho|X} = \tr\p{\rho X}$ for a qudit state $\rho$.
 
@@
   f_{\ell m} = \sum_{\v v} \tilde C_{\ell m}\p{\v v} f\p{\v v}
-  = \sum_{\v v,k,n} \tilde C_{\ell m}\p{\v v} f_{kn} Y_{kn}\p{\v v}.
+  = \sum_{\v v,k,n} \tilde C_{\ell m}\p{\v v} Y_{kn}\p{\v v} f_{kn}.
 \end{align}
-Reconstruction of functions with arbitrary coefficients $f_{\ell m}$ then implies that
+Reconstruction of functions with arbitrary coefficients $f_{\ell m}$ implies that
 \begin{align}
@@
 
-If the function $f$ is {\it band-limited} at degree $L$, which is to say that $f_{\ell m}=0$ for all $\ell\ge L$, then the full recovery problem in Eq.~\eqref{eq:full_recovery} is provably solvable with a suitable choice of $\abs{V}=L^2$ points \cite{freeden2008spherical, freeden2018spherical}.
+If the function $f$ is {\it band-limited} at degree $L$, which is to say that $f_{\ell m}=0$ for all $\ell\ge L$, then the full recovery problem in Eq.~\eqref{eq:full_recovery} is provably solvable with a suitable choice of $\abs{V}=L^2$ points on the sphere \cite{freeden2008spherical, freeden2018spherical}.
 The existence of these solutions to the full recovery problem in turn implies the existence of $d^2$ measurement axes that suffice to reconstruct arbitrary states of $d$-level spin qudit, whose possible states are essentially ``band-limited'' at degree $d$ \footnote{Spin qudit states can be faithfully represented by quasi-probability distributions on a sphere.
 For a $d$-level spin, these quasi-probability distributions are band-limited at degree $d$.}.
-Moreover, for any fixed degree $\ell$, finding solutions to the reduced recovery problem in Eq.~\eqref{eq:tomo_recovery} is equivalent to the recovery of a degree-$\ell$ function $f = \sum_m f_m Y_{\ell m}$, which provably is possible with $\abs{V}=2\ell+1$ samples \cite{freeden2008spherical}.
+Moreover, for any fixed degree $\ell$, finding solutions to the reduced recovery problem in Eq.~\eqref{eq:tomo_recovery} is equivalent to the recovery of a degree-$\ell$ function $f = \sum_m f_m Y_{\ell m}$, which is provably possible with $\abs{V}=2\ell+1$ samples \cite{freeden2008spherical}.
 In the case of spin qudit tomography, the degree $\ell$ takes a maximal value of $\ell_{\t{max}}\equiv d-1$, so state recovery requires at least as many measurement axes as there are transition operators with degree $\ell_{\t{max}}$, namely $2\ell_{\t{max}}+1=2d-1$.
@@
 \item there are only $2\ell+1\le2d-1$ transition operators of degree $\ell\le\ell_{\t{max}}$, and
-\item measuring spin projections along any axis simultaneously provides measurement data for transition operators of {\it all} degrees $\ell\le\ell_{\t{max}}$,
+\item measuring spin projection along any single axis provides measurement data for transition operators of {\it all} degrees $\ell\le\ell_{\t{max}}$,
 \end{enumerate*}
 one might expect that $2d-1$ suitably chosen measurement axes would always suffice to perform full tomography of a $d$-level spin qudit.
-In the language of classical signal recovery, we would expect there to exist a set of $2\ell_{\t{max}}+1$ points on the sphere that enable the recovery of any function with fixed degree $\ell\le\ell_{\t{max}}$ (given prior knowledge of $\ell$) \footnote{We could further conjecture that if {\it any} set of points $V\subset\Omega_2$ enables the recovery of degree-$\ell_{\t{max}}$ functions on the sphere, then $V$ would allow recovery degree-$\ell$ functions with $\ell\le\ell_{\t{max}}$.
+In the language of classical signal recovery, we would expect there to exist a fixed set of $2\ell_{\t{max}}+1$ points on the sphere that enable the recovery of {\it any} function with fixed degree $\ell\le\ell_{\t{max}}$ (given prior knowledge of $\ell$) \footnote{We could further conjecture that if {\it any} set of points $V\subset\Omega_2$ enables the recovery of degree-$\ell_{\t{max}}$ functions on the sphere, then $V$ would allow recovery degree-$\ell$ functions with $\ell\le\ell_{\t{max}}$.
 However, this condition is stronger than that necessary to guarantee full recovery of a $d$-level spin qudit state with $2d-1$ measurement axes.}.
@@
 For the practically minded, the mere existence of solutions to a problem is less interesting than the exposition of a particular solution.
-On a high level, a spin qudit tomography protocol needs to
+On a high level, a spin qudit tomography protocol consists of
 \begin{enumerate*}
-\item select a set of measurement axes,
-\item collect measurement data on spin projection onto these axes, and then
+\item selecting a set of measurement axes,
+\item collecting measurement data on spin projection onto these axes, and then
   \label{tomo:measure}
-\item process the collected data to reconstruct the state of the spin qudit.
+\item processing the collected data to reconstruct the state of the spin qudit.
 \end{enumerate*}
@@
 Intuitively, a good choice of axes should minimize the error with which one can reconstruct an unknown quantum state from associated measurement data.
-If we can quantify this intuition, then we can try to optimize over different choices of measurement axes to find a set that (approximately) minimizes the error in reconstructed states.
+If we can quantify this intuition, then we can optimize over different choices of measurement axes to find a set that (approximately) minimizes the error in reconstructed states.
 
@@
 In this case $M_V$ must be full rank, with $d^2$ nonzero singular values.
-Indexing these singular values $M^V_k$ and the corresponding (normalized) right singular vectors $\v x^V_k\equiv(x^V_{k,1},x^V_{k,2},\cdots)$ by an integer $k\in\set{1,2,\cdots,d^2}$, we can construct the orthonormal qudit operators
+Indexing these singular values $M^V_k$ and the corresponding (normalized) right singular vectors $\v x^V_k\equiv(x^V_{k,1},x^V_{k,2},\cdots,x^V_{k,\abs{V}})$ by an integer $k\in\set{1,2,\cdots,d^2}$, we can construct the orthonormal qudit operators
 \begin{align}
@@
 \end{align}
-The singular values of $M_V$ allow us to make concrete statements about the statistical error between the empirical estimate $\tilde\rho_V$ and the true state $\rho$.
+The measurement matrix $M_V$ allows us to make concrete statements about the statistical error between the empirical estimate $\tilde\rho_V$ and the true state $\rho$.
 Assume, for example, that the estimates $\tilde T_j$ are equal to $\bk{T_j}_\rho$ up to uncorrelated noise with variance at most $\epsilon^2$:
@@
 \end{align}
-Using the fact that the operators $Q^V_k$ are orthonormal, we can therefore bound the mean squared (Euclidean) distance (i.e.~the distance induced by the trace inner product) between $\tilde\rho_V$ and $\rho$ as
+Using the fact that the operators $Q^V_k$ are orthonormal, we can therefore bound the mean squared (Euclidean) distance between $\tilde\rho_V$ and $\rho$ as
 \begin{align}
   \Bbk{\norm{\tilde\rho_V-\rho}^2}
-  = \sum_k \Bbk{\abs{\tilde\rho^V_k - \rho^V_k}^2}
   \le \epsilon^2 \S_V^2,
@@
 \end{align}
-where
+where $\norm{X} \equiv \sqrt{\obk{X|X}} = \sqrt{\tr\p{X^\dag X}}$ is the (Hilbert-Schmidt, or Frobenius) norm of $X$, and
 \begin{align}
-  \S_V^2 \equiv \obk{M_V^{-1}|M_V^{-1}} = \sum_k\f1{\p{M^V_k}^2},
+  \S_V^2 \equiv \sum_k\f1{\p{M^V_k}^2} = \norm{M_V^{-1}}^2,
 \end{align}
-is the squared (Hilbert-Schmidt, or Frobenius) norm of the right inverse $M_V^{-1}$ of $M_V$.
+is the squared norm of the right inverse $M_V^{-1}$ of $M_V$.
 Note that $\S_V^2$ diverges if $M_V$ is singular, in which case measuring spin projections along all axes in $V$ does not provide sufficient information to reconstruct arbitrary quantum states.
@@
 Computing $\S_V^2$ to certify the bound in Eq.~\eqref{eq:bound_eps} requires constructing the measurement matrix $M_V$ and finding its singular values.
-The complexity of this task can be greatly reduced by the fact that the degree $\ell$ of the transition operator $T_{\ell,0}$ is preserved under rotations, which implies that the unitary
+The complexity of this task can be greatly reduced by the fact that the degree $\ell$ of a transition operator $T_{\ell m}$ is preserved under rotations, which implies that the unitary
 \begin{align}
@@
   &&
-  \S_{V\ell}^2 \equiv \obk{M_{V\ell}^{-1}|M_{V\ell}^{-1}},
+  \S_{V\ell}^2 \equiv \norm{M_{V\ell}^{-1}}^2,
 \end{align}
@@
 If $\abs{V}$ is too large for such optimization  to be practical, one can instead generate many random sets of measurement axes, and then choose the set with the smallest error scale $\epsilon_V^2$.
-After choosing a set of measurement axes $V$, experimentally characterize the expectation values $\bk{\Pi_{\v v\mu}}_\rho$ by making measuring spin projection along every axis $\v v\in V$.
-These spin projections determine the expectation values $\bk{T_{\v v\ell,0}}_\rho$, whose estimates can be used to construct the state $\tilde\rho_V$ defined in Eq.~\eqref{eq:reconstructed_state}.
-If spin projection was measured $n$ times along every axis, then the reconstructed state $\tilde\rho_V$ approximates the true state $\rho$ to within a mean squared (Euclidean) distance $\epsilon_V^2/n$.
-Furthermore, if $\tilde\rho_V$ has negative eigenvalues, its distance from $\rho$ can be further reduced with maximum-likelihood corrections \cite{smolin2012efficient}.
+After choosing a set of measurement axes $V$, (experimentally) make $n$ measurements of spin projection along every axis $\v v\in V$, and set $\tilde\Pi_{\v v\mu}\approx\bk{\Pi_{\v v\mu}}_\rho$ to the fraction of times in which the measurement outcome was $\mu$.
+The estimates $\tilde\Pi_{\v v\mu}$ of $\bk{\Pi_{\v v\mu}}_\rho$ determine estimates of the expectation values $\bk{T_{\v v\ell,0}}_\rho$, which can be used to construct the state $\tilde\rho_V$ defined in Eq.~\eqref{eq:reconstructed_state}.
+If $\tilde\rho_V$ has negative eigenvalues, its (Euclidean) distance from $\rho$ can be reduced with maximum-likelihood corrections \cite{smolin2012efficient}.
 
@@
 \end{align}
-is the variance of the entries of $X$ when sampled according to the (classical) probability distribution $p$.
+is the variance of the entries in $X$ when sampled according to the probability distribution $p$.
 The variance $\sigma_p^2\p{X}$ is maximal when $p$ has equal weight on the largest and smallest values of $X$, which implies that
@@
   = \abs{\tilde{\v q}_{\ell i}}^2
-  = \bk{\v v_i | \p{\Sigma_{V\ell}^{-1} W_{V\ell}^\dag}^\dag
-    \p{\Sigma_{V\ell}^{-1} W_{V\ell}^\dag} | \v v_i}
   = \bk{\v v_i | W_{V\ell} \Sigma_{V\ell}^{-2} W_{V\ell}^\dag | \v v_i}

### 2020-11-12 sun_phases/spin_qudit_tomo.tex
find maximizer of the reconstruction error (in appendix)

@@
 
+\newcommand{\C}{\mathcal{C}}
+\newcommand{\D}{\mathcal{D}}
+\newcommand{\I}{\mathcal{I}}
+\renewcommand{\L}{\mathcal{L}}
+\newcommand{\N}{\mathcal{N}}
 \renewcommand{\O}{\mathcal{O}}
@@
 \begin{align}
-  \rho = \sum_{\ell=0}^{d-1} \sum_{m=-\ell}^\ell \rho_{\ell m} T_{\ell m},
+  \rho = \sum_{\ell=0}^{d-1} \sum_{m=-\ell}^\ell
+  \rho_{\ell m} T_{\ell m}^\dag,
   &&
-  \rho_{\ell m} \equiv \obk{T_{\ell m}|\rho},
+  \rho_{\ell m} \equiv \bk{T_{\ell m}}_\rho,
   \label{eq:trans_state}
 \end{align}
-where $\rho^\dag=\rho$ implies that $\rho_{\ell m}^*=\p{-1}^m\rho_{\ell,-m}$.
+where $\bk{X}_\rho\equiv\obk{\rho|X}=\tr\p{\rho X}$, and $\rho^\dag=\rho$ implies that $\rho_{\ell m}^*=\p{-1}^m\rho_{\ell,-m}$.
 The transition operators can be physically interpreted as the transition $\ket\psi\to T_{\ell m}\ket\psi$ induced on the state $\ket\psi$ of a spin-$s$ system by the absorption of a spin-$\ell$ particle with spin projection $m$ onto a fixed quantization axis.
@@
 
-Our goal is to reconstruct an arbitrary state of a spin qudit from measurements of spin projection onto different quantization axes.
+Our goal is to reconstruct an arbitrary state $\rho$ of a spin qudit from measurements of spin projection onto different quantization axes.
 We are thus nominally restricted to measuring projectors $\Pi_{\v v\mu} \equiv \op{\mu_{\v v}}$, where $\ket{\mu_{\v v}}\equiv R\p{\v v}\ket{\mu}$ is a state with spin projection $\mu$ onto a measurement axis $\v v$ determined by a point on the sphere.
 For any fixed axis $\v v$, the sets $\set{\Pi_{\v v\mu}}$ and $\set{T_{\v v\ell,0}}$ are both orthonormal bases for the space of operators that are diagonal in the basis $\set{\ket{\mu_{\v v}}}$.
-By a simple change of operator basis, measuring the projectors $\set{\Pi_{\v v\mu}}$ is therefore equivalent to measuring the transition operators $\set{T_{\v v\ell,0}}$, and provides data on the expectation values $\bk{T_{\v v\ell,0}}_\rho$, where $\bk{X}_\rho \equiv \obk{\rho|X} = \tr\p{\rho X}$ for a qudit state $\rho$.
+By a simple change of operator basis, measuring the projectors $\set{\Pi_{\v v\mu}}$ is therefore equivalent to measuring the transition operators $\set{T_{\v v\ell,0}}$, and provides data on the expectation values $\bk{T_{\v v\ell,0}}_\rho$.
 
@@
 \begin{align}
-  \rho_{\ell m}^* = \bk{T_{\ell m}}_\rho
+  \rho_{\ell m} = \bk{T_{\ell m}}_\rho
   = \sum_{\v v,k} C_{\ell mk}\p{\v v} \bk{T_{\v vk,0}}_\rho.
@@
 \begin{align}
-  \rho = \sum_{k=1}^{d^2} \rho_k^V Q_k^V,
+  \rho = \sum_{k=1}^{d^2} \rho_k^V \p{Q_k^V}^\dag,
   &&
@@
 \begin{align}
-  \tilde\rho_V \equiv \sum_k \tilde\rho^V_k Q^V_k,
+  \tilde\rho_V \equiv \sum_k \tilde\rho^V_k \p{Q^V_k}^\dag,
   \label{eq:reconstructed_state}
@@
 \begin{align}
-  \xi_{V\ell} &\equiv M_{V\ell}
+  \N_{V\ell} &\equiv M_{V\ell}
   \diag\sp{M_{V\ell}^{-1} \p{M_{V\ell}^{-1}}^\dag} M_{V\ell}^\dag, \\
@@
 where $\diag\sp{X}$ sets all off-diagonal entries of $X$ to zero, and $\cov_\rho\p{X,Y} \equiv \bk{XY}_\rho - \bk{X}_\rho \bk{Y}_\rho$.
-We discuss details of efficiently computing $\xi_{V\ell}$ in Appendix \ref{sec:compute}.
+We discuss details of efficiently computing $\N_{V\ell}$ in Appendix \ref{sec:compute}.
 These matrices allow us to compactly express
@@
   \Bbk{\norm{\tilde\rho_V-\rho}^2}
-  = \f1n \sum_\ell \p{\xi_{V\ell}|\T_{\rho\ell}}.
+  = \f1n \sum_\ell \p{\N_{V\ell}|\T_{\rho\ell}}.
   \label{eq:error}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Relaxing assumptions of the reconstruction error bound}
+\section{Revisiting the reconstruction error bound}
 \label{sec:error}
@@
   \tilde \Pi_{i\mu} \equiv \f1n \sum_{j=1}^n \tilde \Pi_{i\mu}^j
-  \approx \f1n \sum_{j=1}^n \tr\p{\Pi_{i\mu}^j \rho^{\otimes N}}
-  = \tr\p{\Pi_{i\mu} \rho}.
+  \approx \f1n \sum_{j=1}^n \tr\p{\rho^{\otimes N} \Pi_{i\mu}^j}
+  = \tr\p{\rho \Pi_{i\mu}}.
 \end{align}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Covariance of errors in the spin-projection basis}
@@
     - \bk{\Pi_{i\mu}}_\rho \bk{\Pi_{i\mu'}}_\rho} \\
-  &= \delta_{ii'} \times \f1n \sp{\delta_{\mu\mu'} p_\mu^i - p_\mu^i p_{\mu'}^i},
+  &= \delta_{ii'} \times \f1n \cov_\rho\p{\Pi_{i\mu}, \Pi_{i\mu'}},
 \end{align}
-where $p_\mu^i \equiv \bk{\Pi_{i\mu}}_\rho = \bk{\mu_{\v v_i}|\rho|\mu_{\v v_i}}$ are the diagonal entries of $\rho$ in the basis of spin projections along axis $\v v_i$.
+where $\cov_\rho\p{X,Y} \equiv \bk{XY}_\rho - \bk{X}_\rho \bk{Y}_\rho$.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Covariance of errors in the transition operator basis}
@@
   \bbk{\epsilon_{i\mu} \epsilon_{i'\mu'}}
-  = \delta_{ii'} \times \f1n \sum_{\mu,\mu'} t_{\ell\mu} t_{\ell'\mu'}
-  \p{\delta_{\mu\mu'} p_\mu^i - p_\mu^i p_{\mu'}^i}
+  = \delta_{ii'} \times \f1n
+  \sum_{\mu,\mu'} t_{\ell\mu} t_{\ell'\mu'}
+  \cov_\rho\p{\Pi_{i\mu}, \Pi_{i\mu'}}
   = \delta_{ii'} \times \f1n \cov_\rho\p{T_{i\ell}, T_{i\ell'}},
 \end{align}
-where $\cov_\rho\p{X,Y} \equiv \bk{XY}_\rho - \bk{X}_\rho \bk{Y}_\rho$.
-In order to establish an upper bound on reconstruction error for an unknown quantum state, we will need an upper bound on the variance $\bbk{\epsilon_{i\ell}^2}$ in particular.
-To this end, we observe that
+where we used the fact that the covariance $\cov_\rho\p{X,Y}$ is linear in both $X$ and $Y$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Exact reconstruction error}
+
+We now revisit the derivation of reconstruction error in Section \ref{sec:bound}.
+To summarize, for a set of measurement axes $V=\set{\v v}$ and degrees $\ell\in\set{0,1,\cdots,d-1}$ we construct the measurement matrix
+\begin{align}
+  M_V \equiv \sum_{\v v,\ell} \oket{T_{\v v\ell,0}} \bra{\v v\ell},
+\end{align}
+which is block diagonalized as
+\begin{align}
+  U \equiv \sum_{\ell,m} \ket{\ell m} \obra{T_{\ell m}},
+  &&
+  U M_V = \sum_\ell \op{\ell} \otimes M_{V\ell},
+  &&
+  M_{V\ell} = \sum_{m,\v v} D^\ell_{m,0}\p{\v v} \op{m}{\v v},
+\end{align}
+where $D^\ell_{mn}\p{\v v}\equiv\bk{\ell m|R\p{\v v}|\ell n}$ is a (Wigner) rotation matrix element for a spin-$\ell$ particle.
+The block-diagonal structure of $M_V$ allows us to index its singular values $M^V_{\ell m}$ and corresponding (normalized) right singular vectors $\v x^V_{\ell m} = (x^V_{\ell m,1}, x^V_{\ell m,2}, \cdots )$ by the indices $\p{\ell,m}$, where the integer $\abs{m}\le\ell$.
+These singular vectors and values allow us to define the orthonormal operators
+\begin{align}
+  Q^V_{\ell m} \equiv \sum_i q^V_{\ell mi} T_{i\ell},
+  &&
+  \v q^V_{\ell m} \equiv \f{\v x^V_{\ell m}}{M^V_{\ell m}},
+\end{align}
+where $i\in\set{1,2,\cdots,\abs{V}}$ indexes an axis $\v v_i\in V$.
+The state $\rho$ can be expanded in the basis of $\set{Q^V_{\ell m}}$ as
+\begin{align}
+  \rho = \sum_{\ell,m} \bk{Q^V_{\ell m}}_\rho \p{Q^V_{\ell m}}^\dag,
+\end{align}
+and the estimates $\tilde T_{i\ell}$ of $\bk{T_{i\ell}}_\rho$ can be used to construct the following estimate $\tilde\rho_V$ of $\rho$:
+\begin{align}
+  \tilde\rho_V
+  \equiv \sum_{\ell,m} \sp{\sum_i q^V_{\ell mi} \tilde T_{i\ell}}
+  \p{Q^V_{\ell m}}^\dag
+  \approx \sum_{\ell,m} \sp{\sum_i q^V_{\ell mi} \bk{T_{i\ell}}_\rho}
+  \p{Q^V_{\ell m}}^\dag
+  = \sum_{\ell,m} \bk{Q^V_{\ell m}}_\rho \p{Q^V_{\ell m}}^\dag
+  = \rho.
+\end{align}
+Expanding $\tilde T_{i\ell} = \bk{T_{i\ell}}_\rho + \epsilon_{i\ell}$, we can use orthonormality of all $Q^V_{\ell m}$ to expand the mean squared distance between $\tilde\rho_V$ and $\rho$ as
+\begin{align}
+  \Bbk{\norm{\tilde\rho_V-\rho}^2}
+  = \sum_{\ell,m,i,i'} \p{q^V_{\ell mi}}^* q^V_{\ell mi'} \bbk{\epsilon_{i\ell} \epsilon_{i'\ell}}
+  = \f1n \sum_{\ell,i} \abs{\tilde{\v q}_{\ell i}}^2 \cov_\rho\p{T_{i\ell},T_{i\ell}},
+  &&
+  \abs{\tilde{\v q}_{\ell i}}^2 = \sum_m \abs{q_{\ell mi}}^2.
+  \label{eq:error_SM_start}
+\end{align}
+where $\tilde{\v q}_{\ell i} = (q_{\ell,-\ell,i},q_{\ell,-\ell+1,i},\cdots q_{\ell,\ell,i})$.
+Identifying the singular value decomposition $M_{V\ell} = U_{V\ell} \Sigma_{V\ell} W_{V\ell}^\dag$, the we observe that vector $\tilde{\v q}_{\ell i}$ is essentially the $i$-th row of $\Sigma_{V\ell}^{-1} W_{V\ell}^\dag$, which is to say that $\tilde{\v q}_{\ell i} = \Sigma_{V\ell}^{-1} W_{V\ell}^\dag \ket{\v v_i}$.
+We can therefore simplify
+\begin{align}
+  \abs{\tilde{\v q}_{\ell i}}^2
+  = \bk{\v v_i | W_{V\ell} \Sigma_{V\ell}^{-2} W_{V\ell}^\dag | \v v_i}
+  = \bk{\v v_i| M_{V\ell}^{-1} \p{M_{V\ell}^{-1}}^\dag | \v v_i}.
+\end{align}
+Using the fact that all $T_{i\ell}=T_{i\ell}^\dag$, we can also expand
+\begin{align}
+  \cov_\rho\p{T_{i\ell},T_{i\ell}}
+  = \cov_\rho\p{T_{i\ell}^\dag,T_{i\ell}}
+  = \sum_{m,n} D^\ell_{m,0}\p{\v v_i}^* D^\ell_{n,0}\p{\v v_i}
+  \cov_\rho\p{T_{\ell m}^\dag, T_{\ell n}},
+\end{align}
+which implies that
+\begin{align}
+  \Bbk{\norm{\tilde\rho_V-\rho}^2}
+  = \f1n \sum_{i,\ell,m,n}
+  \abs{\tilde{\v q}_{\ell i}}^2
+  D^\ell_{m,0}\p{\v v_i}^* D^\ell_{n,0}\p{\v v_i}
+  \cov_\rho\p{T_{\ell m}^\dag, T_{\ell n}}.
+\end{align}
+Altogether, the exact reconstruction error in Eq.~\eqref{eq:error_SM_start} can be expressed more compactly by defining the covariance matrix
+\begin{align}
+  \T_{\rho\ell} \equiv
+  \sum_{m,n} \cov_\rho\p{T_{\ell m}^\dag, T_{\ell n}} \op{m}{n},
+\end{align}
+and the noise matrix
+\begin{align}
+  \N_{V\ell} \equiv M_{V\ell} \diag\sp{M_{V\ell}^{-1}
+    \p{M_{V\ell}^{-1}}^\dag} M_{V\ell}^\dag,
+\end{align}
+where $\diag\sp{X}$ sets all off-diagonal entries of $X$ to zero, in terms of which
+\begin{align}
+  \Bbk{\norm{\tilde\rho_V-\rho}^2}
+  = \f1n \sum_\ell \obk{\N_{V\ell}|\T_{\rho\ell}}.
+  \label{eq:error_SM}
+\end{align}
+While the error in Eq.~\eqref{eq:error_SM} cannot be known exactly without knowing $\rho$, this error can be estimated by computing the right-hand side of Eq.~\eqref{eq:error_SM} with the replacement $\rho\to\tilde\rho_V$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{An improved reconstruction error bound}
+
+We now seek to maximize the error in Eq.~\eqref{eq:error_SM} over all $\rho$, which will provide an upper bound on reconstruction error.
+To this end, we vary the error with respect to $\rho$ by considering the vector derivative (i.e.~gradient)
+\begin{align}
+  \f{\partial}{\partial\obra\rho} \Bbk{\norm{\tilde\rho_V-\rho}^2}
+  = \f1n \sum_\ell \tr\p{\N_{V\ell}
+    \f{\partial}{\partial\obra\rho} \T_{\rho\ell}}
+  = 0,
+  &&
+  \f{\partial}{\partial\obra\rho}
+  \equiv \sum_{L,M} \oket{T_{LM}} \f{\partial}{\partial\rho_{LM}}.
+  \label{eq:variation}
+\end{align}
+Here $\partial/\partial\obra\rho$ is essentially a vector of all $\partial/\partial\rho_{LM}$, which satisfies
+\begin{align}
+  \f{\partial}{\partial\obra\rho} \obk{\rho|X}
+  = \f{\partial}{\partial\obra\rho}
+  \sum_{\ell,m} \obk{\rho|T_{\ell m}} \obk{T_{\ell m}|X}
+  = \sum_{L,M} \oket{T_{LM}} \f{\partial}{\partial\rho_{LM}}
+  \sum_{\ell,m} \rho_{\ell m} \obk{T_{\ell m}|X}
+  = \sum_{L,M} \oket{T_{LM}} \obk{T_{LM}|X}
+  = \oket{X}.
+\end{align}
+The variation in Eq.~\eqref{eq:variation} leads us to consider
+\begin{align}
+  \f{\partial}{\partial\obra\rho} \T_{\rho\ell}
+  &= \sum_{m,n} \op{m}{n} \otimes \f{\partial}{\partial\obra\rho}
+  \sp{\bk{T_{\ell m}^\dag T_{\ell n}}_\rho
+    - \bk{T_{\ell m}^\dag}_\rho \bk{T_{\ell n}}_\rho} \\
+  &= \sum_{m,n} \op{m}{n} \otimes \f{\partial}{\partial\obra\rho}
+  \sp{\obk{\rho|T_{\ell m}^\dag T_{\ell n}}
+    - \obk{\rho|T_{\ell m}^\dag} \obk{\rho|T_{\ell n}}} \\
+  &= \sum_{m,n} \op{m}{n} \otimes \sp{\oket{T_{\ell m}^\dag T_{\ell n}}
+    - \oket{T_{\ell m}^\dag} \obk{\rho|T_{\ell n}}
+    - \obk{\rho|T_{\ell m}^\dag} \oket{T_{\ell n}}}.
+  \label{eq:variation_cov}
+\end{align}
+The $\sim\rho$ terms in Eq.~\eqref{eq:variation_cov} can be simplified by the fact that
+\begin{align}
+  \obk{\rho|T_{\ell m}}
+  = \tr\p{\rho T_{\ell m}}
+  = \p{-1}^m \tr\p{T_{\ell,-m}^\dag \rho}
+  = \p{-1}^m \obk{T_{\ell,-m}|\rho}
+  = \obk{T_{\ell m}^\dag|\rho},
+\end{align}
+which implies that
+\begin{align}
+  \sum_{m,n} \op{m}{n} \otimes
+  \sp{\oket{T_{\ell m}^\dag} \obk{\rho|T_{\ell n}}
+    + \obk{\rho|T_{\ell m}^\dag} \oket{T_{\ell n}}}
+  &= \sum_{m,n} \op{m}{n} \otimes
+  \sp{\oket{T_{\ell m}^\dag} \obk{T_{\ell n}^\dag|\rho}
+    + \oket{T_{\ell n}} \obk{T_{\ell m}|\rho}} \\
+  &= \sp{\sum_{m,n} \p{\I\op{m}{n}\I + \op{n}{m}}
+    \otimes \oop{T_{\ell m}}{T_{\ell n}}} \oket{\rho},
+\end{align}
+where we define the inversion operator $\I\equiv\sum_m \p{-1}^m\op{-m}{m}$.
+Meanwhile, the constant terms in Eq.~\eqref{eq:variation_cov} are
+\begin{align}
+  \sum_{m,n} \op{m}{n} \otimes \oket{T_{\ell m}^\dag T_{\ell n}}
+  = \sum_{L,m,n} \I \op{m}{n} \otimes g_{\ell mn}^L \oket{T_{L,m+n}},
+  &&
+  g_{\ell mn}^L \equiv \obk{T_{L,m+n}|T_{\ell m} T_{\ell n}}
+  = f_{\ell m;\ell n}^{L,m+n}
+\end{align}
+where the factors $f_{\ell m;\ell n}^{L,m+n}$ are provided in Appendix \ref{sec:trans_prod}.
+Altogether, we have that
+\begin{align}
+  \f{\partial}{\partial\obra\rho} \T_{\rho\ell}
+  = \sum_{m,n} g_{\ell mn}^L \I\op{m}{n} \otimes \oket{T_{L,m+n}}
+  - \sp{\sum_{m,n} \p{\op{n}{m} + \I\op{m}{n}\I}
+    \otimes \oop{T_{\ell m}}{T_{\ell n}}} \oket{\rho}.
+\end{align}
+Substituting this result back into Eq.~\eqref{eq:variation}, we find that
+\begin{align}
+  n\, \f{\partial}{\partial\obra\rho}
+  \Bbk{\norm{\tilde\rho_V-\rho}^2}
+  = \oket{\chi_V} - \sum_\ell \L_{V\ell} \oket\rho
+  = 0,
+\end{align}
+where
+\begin{align}
+  \oket{\chi_V} \equiv \sum_{L,M} \chi^V_{LM} \oket{T_{LM}},
+  &&
+  \L_{V\ell} \equiv \sum_{m,n}
+  \sp{\bk{m|\N_{V\ell}|n} + \bk{n|\I\N_{V\ell}\I|m}}
+  \oop{T_{\ell m}}{T_{\ell n}}.
+\end{align}
+\begin{align}
+  \chi^V_{LM} \equiv \sum_\ell \obk{\N_{V\ell}|\D_M|\I g_{L\ell}},
+  &&
+  g_{L\ell} \equiv \sum_{m,n} g^L_{\ell mn} \op{m}{n},
+  &&
+  \D_M \equiv \sum_{m,n} \delta_{M,m+n} \op{mn}{mn}.
+\end{align}
+Defining the projection of $\oket{\chi_V}$ onto the degree-$\ell$ subspace of transition operators,
+\begin{align}
+  \oket{\chi_{V\ell}}
+  \equiv \sum_m \oket{T_{\ell m}} \obk{T_{\ell m}|\chi_V},
+\end{align}
+we thus find that reconstruction error is maximized by the state
+\begin{align}
+  \oket\rho = \sum_\ell \L_{V\ell}^{-1} \oket{\chi_{V\ell}},
+\end{align}
+where $\L_{V\ell}^{-1}$ is the inverse of $\L_{V\ell}$.
+
+[\red{todo: substitute this solution back into the reconstruction error to find an upper bound}]
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Simpler (and looser) upper bounds}
+
+In order to establish simpler bounds on reconstruction error, we define the probability $p^i_\mu\equiv\bk{\Pi_{i\mu}}_\rho$, collect these probabilities into the classical probability distribution $p^i = \p{p^i_{-s}, p^i_{-s+1}, \cdots, p^i_s}$, and define the vector $t_\ell \equiv \p{t_{\ell,-s},t_{\ell,-s+1},\cdots,t_{\ell s}}$ of all $t_{\ell\mu}\equiv\bk{\mu|T_{\ell,0}|\mu}$.
+We then observe that
 \begin{align}
@@
 \end{align}
-where $t_\ell \equiv \p{t_{\ell,-s},t_{\ell,-s+1},\cdots,t_{\ell s}}$ is a vector of all $t_{\ell\mu}$, and
+where
 \begin{align}
@@
   \equiv \sum_\mu p_\mu X_\mu^2 - \p{\sum_\mu p_\mu X_\mu}^2
-  = \sum_\mu p_\mu \p{X_\mu - \sum_\nu p_\nu X_\nu}^2
 \end{align}
-is the variance of the entries in $X$ when sampled according to the probability distribution $p$.
-The variance $\sigma_p^2\p{X}$ is maximal when $p$ has equal weight on the largest and smallest values of $X$, which implies that
+is the weighted variance of $X$.
+This variance is maximal when $p$ has equal weight on the largest and smallest values of $X$, which implies that
 \begin{align}
@@
   &&
-  \Gamma_\ell \equiv \f{\max_\mu t_{\ell\mu} - \min_\mu t_{\ell\mu}}{2}.
+  \Gamma_\ell \equiv \f{\max_\mu t_{\ell\mu} - \min_\mu t_{\ell\mu}}{2},
 \end{align}
-Altogether, using additionally the fact that $\sigma_p^2\p{t_0}=0$ (because all $t_{0,\mu}=1/\sqrt{d}$) we can bound
+which lets us bound
 \begin{align}
   \bbk{\epsilon_{i\ell}^2}
-  \le \f1n \times \Gamma_\ell^2 \times \p{1 - \delta_{\ell,0}}
-  = \f{2\ell+1}{nd} \times \gamma_\ell^2
-  \times \p{1 - \delta_{\ell,0}},
+  \le \f1n \times \Gamma_\ell^2
+  = \f{2\ell+1}{nd} \times \gamma_\ell^2,
   \label{eq:variance_bound}
@@
 \end{align}
-Note that the bound in Eq.~\eqref{eq:variance_bound} is tight, as equality is achieved by the mixed state
+The bound in Eq.~\eqref{eq:variance_bound} lets us bound reconstruction error as
 \begin{align}
-  \rho^\star
-  \equiv \f12 \p{\Pi_{\v v_i,\mu_{\t{max}}} + \Pi_{\v v_i,\mu_{\t{min}}}},
+  \Bbk{\norm{\tilde\rho_V-\rho}^2}
+  = \sum_{\ell,m,i} \abs{q^V_{\ell mi}}^2 \bbk{\epsilon_{i\ell}^2}
+  \le \f1n \sum_{\ell,m,i} \abs{q^V_{\ell mi}}^2 \Gamma_\ell^2
+  = \f1n \sum_\ell \Gamma_\ell^2 \S_{V\ell}^2
+  = \f1{nd} \sum_\ell \p{2\ell+1} \gamma_\ell^2 \S_{V\ell}^2,
 \end{align}
-where $\mu_{\t{max}}$ ($\mu_{\t{min}}$) is any index that maximizes (minimizes) $\tau_{\ell\mu}$.
-
-Though the result in Eq.~\eqref{eq:variance_bound} is sufficient to proceed with our derivation of an improved tomographic reconstruction error bound, we first take some time to build a better analytical understanding of the bound on $\bbk{\epsilon_{i\ell}^2}$.
-To this end, we first note that the Clebsch-Gordan coefficients $\tau_{\ell\mu}$ are at most $1$ in magnitude, which implies that $\gamma_\ell^2\le 1$.
+where
+\begin{align}
+  \S_{V\ell}^2 = \sum_{m,i} \abs{q_{\ell mi}}^2
+  = \sum_m \f1{\p{M^V_{\ell m}}^2}
+  = \norm{M_{V\ell}^{-1}}^2.
+\end{align}
+To better understand this bound, we note that the Clebsch-Gordan coefficients $\tau_{\ell\mu}$ are at most $1$ in magnitude, which implies that $\gamma_\ell^2\le 1$.
 Moreover, we can also bound
@@
   &&
-  \bbk{\epsilon_{i\ell}^2} \le \f1n \times \min\Set{\f{2\ell+1}{d}, 1}.
+  \Gamma_\ell^2 \le \min\Set{\f{2\ell+1}{d}, 1},
 \end{align}
-We can get a tighter analytical bound on $\bbk{\epsilon_{i\ell}^2}$ by considering the fact that $t_{\ell\mu_{\t{max}}}^2 = t_{\ell,-\mu_{\t{max}}}^2$ due to the symmetries of the Clebsch-Gordan coefficients, so if $\mu_{\t{max}}\ne0$, then
+where $\S_V^2=\sum_\ell\S_{V\ell}^2$.
+We can get a tighter analytical bound on $\Gamma_\ell^2$ by considering the fact that $t_{\ell\mu}^2 = t_{\ell,-\mu}^2$ due to the symmetries of the Clebsch-Gordan coefficients.
+Denoting the index that maximizes $t_{\ell\mu}^2$ by $\mu_{\t{max}}$, it follows that if $\mu_{\t{max}}\ne0$ then
 \begin{align}
@@
   \stackrel{\mu_{\t{max}}=0}{\le}
-  \sqrt{\f{1 - t_{\ell\mu_{\t{max}}}^2}{2}}.
+  \sqrt{\f{1 - t_{\ell\mu_{\t{max}}}^2}{2}},
 \end{align}
+where $\mu_{\t{min}}$ is the index that minimizes $t_{\ell\mu}^2$.
 This inequality lets us bound
@@
   + \f12\sqrt{\f{1 - t_{\ell\mu_{\t{max}}}^2}{2}}
-  \equiv g\p{t_{\ell\mu_{\t{max}}}}.
+  \equiv \lambda\p{t_{\ell\mu_{\t{max}}}}.
 \end{align}
-If is straightforward to show that $g\p{M}$ has a single maximum at $M=\sqrt{2/3}$, and that $g^\star\equiv g\p{\sqrt{2/3}}=\sqrt{3/8}$, so
+If is straightforward to show that $\lambda\p{x}$ has a single maximum at $x=\sqrt{2/3}$, and that $\lambda^\star\equiv \lambda\p{\sqrt{2/3}}=\sqrt{3/8}$, so
 \begin{align}
-  \sigma_p^2\p{t_\ell} \le \Gamma_\ell^2
-  \stackrel{\mu_{\t{max}}=0}{\le} \p{g^\star}^2
+  \Gamma_\ell^2
+  \stackrel{\mu_{\t{max}}=0}{\le} \p{\lambda^\star}^2
   = \f38 < \f12.
@@
 \begin{align}
-  \bbk{\epsilon_{i\ell}^2}
-  \le \f1n \times \min\Set{\f{2\ell+1}{d}, \f12}.
+  \Gamma_\ell^2 \le \min\Set{\f{2\ell+1}{d}, \f12},
+  &&
+  \t{so}
+  &&
+  \Bbk{\norm{\tilde\rho_V-\rho}^2}
+  \le \f1n \sum_\ell \Gamma_\ell^2 \S_{V\ell}^2
+  \le \f{\S_V^2}{2n},
 \end{align}
-
-\subsection{An improved reconstruction error bound}
-
-Equipped with the bound on variances in Eq.~\eqref{eq:variance_bound}, we again consider the block-diagonalized form of the measurement matrix derived in Section \ref{sec:bound}:
-\begin{align}
-  U M_V = \sum_\ell \op{\ell} \otimes M_{V\ell},
-  &&
-  M_{V\ell} = \sum_{m,\v v} D^\ell_{m,0} \op{m}{\v v}.
-\end{align}
-The block diagonal structure of this matrix implies that, rather than an integer index $k\in\set{1,2,\cdots,d^2}$, we can identify the singular values and right singular vectors of $M_V$ by a block index $\ell\in\set{0,1,\cdots,d-1}$ and a sub-index $m=\set{-\ell,-\ell+1,\cdots,\ell}$.
-Making the replacement $k\to\p{\ell,m}$, we can thus repeat the derivations through Eq.~\eqref{eq:bound_eps} of Section \ref{sec:bound} without the assumption of Eq.~\eqref{eq:error_assumption}, finding that
-\begin{align}
-  \Bbk{\abs{\tilde\rho_{\ell m}^V-\rho_{\ell m}^V}^2}
-  = \sum_{i,i'} q_{\ell m i}^V \p{q_{\ell m i'}^V}^*
-  \bbk{\epsilon_{i\ell} \epsilon_{i'\ell}}
-  = \sum_i \abs{q_{\ell m i}^V}^2
-  \bbk{\epsilon_{i\ell}^2}
-  < \f1{\p{M^V_{\ell m}}^2} \times \f{2\ell+1}{nd} \times \gamma_\ell^2
-  \times \p{1-\delta_{\ell,0}},
-  \label{eq:bound_single}
-\end{align}
-where $i$ indexes an axis $\v v_i\in V$.
-Altogether, we thus find that $n$ measurements of spin projection along every axis $\v v\in V$ allows us to construct an empirical estimate $\tilde\rho_V$ of an unknown state $\rho$ that is accurate to within a mean squared distance
-\begin{align}
-  \Bbk{\norm{\tilde\rho_V-\rho}^2}
-  = \sum_{\ell,m} \Bbk{\abs{\tilde\rho_{\ell m}^V-\rho_{\ell m}^V}^2}
-  < \f1{nd} \sum_{\ell>0} \p{2\ell+1} \gamma_\ell^2 \S_{V\ell}^2.
-  \label{eq:bound_SM}
-\end{align}
-We note that whereas the bound on $\bbk{\epsilon_{i\ell}^2}$ in Eq.~\eqref{eq:variance_bound} is tight, the bounds in Eq.~\eqref{eq:bound_single} and Eq.~\eqref{eq:bound_SM} are not: the upper bound on $\bbk{\epsilon_{i\ell}^2}$ cannot be achieved simultaneously for all axes $\v v_i$ or degrees $\ell$.
-In principle, there is therefore room for improvement on the bound in Eq.~\eqref{eq:bound_SM}.
-
-As a final comment, we note that the statistical error bound in \eqref{eq:bound_SM} is a ``worst-case'' bound that is agnostic to the state $\rho$.
-The actual statistical error for any particular state may be considerably smaller, and can be obtained by substituting the exact expression for $\bbk{\epsilon_{i\ell}^2}$:
-\begin{align}
-  \Bbk{\norm{\tilde\rho_V-\rho}^2}
-  = \sum_{i,\ell,m} \abs{q^V_{\ell mi}}^2 \bbk{\epsilon_{i\ell}^2}
-  = \f1n \sum_{\ell,i} C^V_{\ell i} \cov_\rho\p{T_{i\ell},T_{i\ell}},
-  &&
-  C^V_{\ell i} \equiv \sum_m \abs{q_{\ell mi}}^2.
-  \label{eq:error_SM_start}
-\end{align}
-Identifying the singular value decomposition $M_{V\ell} = U_{V\ell} \Sigma_{V\ell} W_{V\ell}^\dag$, the vector $\tilde{\v q}_{\ell i} = (q_{\ell,-\ell,i},q_{\ell,-\ell+1,i},\cdots q_{\ell,\ell,i})$ is essentially the $i$-th row of $\Sigma_{V\ell}^{-1} W_{V\ell}^\dag$, which is to say that $\tilde{\v q}_{\ell i} = \Sigma_{V\ell}^{-1} W_{V\ell}^\dag \ket{\v v_i}$.
-We can therefore simplify
-\begin{align}
-  C^V_{\ell i}
-  = \abs{\tilde{\v q}_{\ell i}}^2
-  = \bk{\v v_i | W_{V\ell} \Sigma_{V\ell}^{-2} W_{V\ell}^\dag | \v v_i}
-  = \bk{\v v_i| M_{V\ell}^{-1} \p{M_{V\ell}^{-1}}^\dag | \v v_i}.
-\end{align}
-Using the fact that all $T_{i\ell}=T_{i\ell}^\dag$, we can also expand
-\begin{align}
-  \cov_\rho\p{T_{i\ell},T_{i\ell}}
-  = \cov_\rho\p{T_{i\ell}^\dag,T_{i\ell}}
-  = \sum_{m,n} D^\ell_{m,0}\p{\v v_i}^* D^\ell_{n,0}\p{\v v_i}
-  \cov_\rho\p{T_{\ell m}^\dag, T_{\ell n}},
-\end{align}
-which implies that
-\begin{align}
-  \Bbk{\norm{\tilde\rho_V-\rho}^2}
-  = \f1n \sum_{i,\ell,m,n}
-  C^V_{\ell i} D^\ell_{m,0}\p{\v v_i}^* D^\ell_{n,0}\p{\v v_i}
-  \cov_\rho\p{T_{\ell m}^\dag, T_{\ell n}}.
-\end{align}
-Altogether, the result in Eq.~\eqref{eq:error_SM_start} can be expressed more compactly using the $\p{2\ell+1}\times\p{2\ell+1}$ matrices
-\begin{align}
-  \xi_{V\ell} \equiv
-  M_{V\ell} \diag\sp{M_{V\ell}^{-1} \p{M_{V\ell}^{-1}}^\dag} M_{V\ell}^\dag,
-  &&
-  \T_{\rho\ell} \equiv
-  \sum_{m,n} \cov_\rho\p{T_{\ell m}^\dag, T_{\ell n}} \op{m}{n},
-\end{align}
-where $\diag\sp{X}$ sets all off-diagonal entries of $X$ to zero, in terms of which
-\begin{align}
-  \Bbk{\norm{\tilde\rho_V-\rho}^2}
-  = \f1n \sum_\ell \p{\xi_{V\ell}|\T_{\rho\ell}}.
-  \label{eq:error_SM}
-\end{align}
-While the error in Eq.~\eqref{eq:error_SM} cannot be known exactly without knowing $\rho$, this error can be estimated by computing the right-hand side of Eq.~\eqref{eq:error_SM} with the replacement $\rho\to\tilde\rho_V$.
+where $\S_V^2 = \sum_\ell \S_{V\ell}^2$.
 

### 2020-11-13 sun_phases/spin_qudit_tomo.tex
minor changes to appendix on reconstruction error, and changes to the corresponding section in the main text

@@
 \renewcommand{\S}{\mathcal{S}}
-\newcommand{\T}{\mathcal{T}}
 
@@
 However, this assumption breaks down when measurement error is limited by fundamental quantum (shot) noise.
-We relax the assumption of Eq.~\eqref{eq:error_assumption} and derive more fundamental error bounds in Appendix \ref{sec:error}.
-Therein, we find that if measurement error is dominated by shot noise, then the estimate $\tilde\rho_V$ built from $n$ independent measurements of spin projection along every axis $\v v\in V$ approximates the true state $\rho$ to within a mean squared distance $\epsilon_V^2/n$, i.e.
-\begin{align}
-  \Bbk{\norm{\tilde\rho_V-\rho}^2} < \f{\epsilon_V^2}{n},
+We relax the assumption of Eq.~\eqref{eq:error_assumption} in Appendix \ref{sec:error}, where we instead assume that $\tilde\rho_V$ is built from $n$ independent measurements of spin projection along every axis $\v v\in V$.
+In this case, the mean squared distance between the approximation $\tilde\rho_V$ and the true state $\rho$ can be compactly expressed as as an inner product of the {\it covariance matrices}
+\begin{align}
+  \C_\ell\sp{\rho} \equiv \sum_{m,n}
+  \cov_\rho\p{T_{\ell m}^\dag, T_{\ell n}} \op{m}{n}
+\end{align}
+where $\cov_\rho\p{X,Y} \equiv \bk{XY}_\rho - \bk{X}_\rho \bk{Y}_\rho$, and the {\it shot noise matrices}
+\begin{align}
+  \N_{V\ell} \equiv M_{V\ell} \diag\sp{M_{V\ell}^{-1}
+    \p{M_{V\ell}^{-1}}^\dag} M_{V\ell}^\dag,
+\end{align}
+where $\diag\sp{X}$ sets all off-diagonal parts of $X$ to zero:
+\begin{align}
+  \Bbk{\norm{\tilde\rho_V-\rho}^2}
+  = \f1n \sum_\ell \obk{\N_{V\ell}|\C_\ell\sp{\rho}}.
+  \label{eq:error}
+\end{align}
+This reconstruction error can be maximized analytically, yielding a reconstruction error bound of the form
+\begin{align}
+  \Bbk{\norm{\tilde\rho_V-\rho}^2} \le \f{\epsilon_V^2}{n},
   \label{eq:bound}
 \end{align}
-where the {\it error scale} $\epsilon_V^2$ is
-\begin{align}
-  \epsilon_V^2 \equiv
-  \f1d \sum_{\ell>0} \p{2\ell+1} \gamma_\ell^2 \S_{V\ell}^2,
-  \label{eq:scale}
-\end{align}
-and $\gamma_\ell$ is defined using Clebsch-Gordan coefficients:
-\begin{align}
-  \gamma_\ell
-  \equiv \f{\max_\mu\tau_{\ell\mu} - \min_\mu\tau_{\ell\mu}}{2},
-  &&
-  \tau_{\ell\mu} \equiv \bk{s\mu;\ell,0|s\mu}.
-\end{align}
-If $d$ even or $\ell$ is odd, then $\gamma_\ell=\max_\mu\tau_{\ell\mu}$.
-For comparison with the previous bound in Eq.~\eqref{eq:bound_eps}, the error scale $\epsilon_V^2<\S_V^2/2$, so the previous bound still holds with the replacement $\epsilon^2\to1/2n$.
-We note that the bound in Eq.~\eqref{eq:bound} is not tight, as it is acquired by bounding the statistical variance $\epsilon_{\v v\ell}^2$ in the empirical estimate of $\bk{T_{\v v\ell,0}}_\rho$ as $\bbk{\epsilon_{\v v\ell}^2}\le\frac{2\ell+1}{nd}\times\gamma_\ell^2$.
-Though the individual bounds for each axis $\v v$ and degree $\ell$
-are tight, these bounds cannot all be achieved simultaneously.
-There is therefore still room for improvement on the bound in Eq.~\eqref{eq:bound}.
-
-As a final comment, we note that Eq.~\eqref{eq:bound} provides a pessimistic upper bound on statistical error, which can be calculated without prior knowledge of the true qudit state $\rho$.
-The actual error in the reconstruction $\tilde\rho_V$ of a particular state $\rho$ may be considerably smaller.
-To write out this error in full (see Appendix \ref{sec:error} for a derivation), we define the $\p{2\ell+1}\times\p{2\ell+1}$ matrices
-\begin{align}
-  \N_{V\ell} &\equiv M_{V\ell}
-  \diag\sp{M_{V\ell}^{-1} \p{M_{V\ell}^{-1}}^\dag} M_{V\ell}^\dag, \\
-  \T_{\rho\ell} &\equiv \sum_{m,n}
-  \cov_\rho\p{T_{\ell m}^\dag, T_{\ell n}} \op{m}{n},
-\end{align}
-where $\diag\sp{X}$ sets all off-diagonal entries of $X$ to zero, and $\cov_\rho\p{X,Y} \equiv \bk{XY}_\rho - \bk{X}_\rho \bk{Y}_\rho$.
-We discuss details of efficiently computing $\N_{V\ell}$ in Appendix \ref{sec:compute}.
-These matrices allow us to compactly express
-\begin{align}
-  \Bbk{\norm{\tilde\rho_V-\rho}^2}
-  = \f1n \sum_\ell \p{\N_{V\ell}|\T_{\rho\ell}}.
-  \label{eq:error}
-\end{align}
-While the error in Eq.~\eqref{eq:error} cannot be known exactly without knowing $\rho$, this error can be estimated by computing the right-hand side of Eq.~\eqref{eq:error} with the replacement $\rho\to\tilde\rho_V$.
-
-[\red{I think the error in Eq.~\eqref{eq:error} can be maximized with no added cost to computational complexity.
-  If so, the second half of this section entire section (as well as Appendix \ref{sec:error}) should be rewritten to replace Eq.~\eqref{eq:bound} by the tighter bound obtained by maximizing Eq.~\eqref{eq:error}.}]
+where the complexity of computing the {\it error scale} $\epsilon_V^2$ is the same as that of computing $\S_V^2$.
+For comparison with the bound in Eq.~\eqref{eq:bound_eps}, we note that the error scale $\epsilon_V^2<\S_V^2/2$, so the previous bound still holds with the replacement $\epsilon^2\to1/2n$.
+
+[\red{todo: finish discussion about $\epsilon_V^2$}]
+
+As a final comment, we note that Eq.~\eqref{eq:bound} only provides a pessimistic upper bound on statistical error, which can be calculated without prior knowledge of the true qudit state $\rho$.
+The actual error in the reconstruction $\tilde\rho_V$ of a particular state $\rho$, provided by Eq.~\eqref{eq:error}, may be considerably smaller.
+While the true error in $\tilde\rho_V$ cannot be known exactly without knowing $\rho$, this error can be estimated by computing the right-hand side of Eq.~\eqref{eq:error} with the replacement $\rho\to\tilde\rho_V$.
 
@@
 \begin{align}
-  \T_{\rho\ell} \equiv
+  \C_\ell\sp{\rho} \equiv
   \sum_{m,n} \cov_\rho\p{T_{\ell m}^\dag, T_{\ell n}} \op{m}{n},
@@
   \Bbk{\norm{\tilde\rho_V-\rho}^2}
-  = \f1n \sum_\ell \obk{\N_{V\ell}|\T_{\rho\ell}}.
+  = \f1n \sum_\ell \obk{\N_{V\ell}|\C_\ell\sp{\rho}}.
   \label{eq:error_SM}
 \end{align}
+In practice, to evaluate this error it is useful to expand the covariance matrix as
+\begin{align}
+  \C_\ell\sp{\rho}
+  = \sum_{m,n} \op{m}{n} \sp{\obk{\rho|T_{\ell m}^\dag T_{\ell n}}
+    - \obk{\rho|T_{\ell m}^\dag} \obk{\rho|T_{\ell n}}}
+  = \sum_{m,n} \I\op{m}{n} \sp{\obk{\rho|T_{\ell m} T_{\ell n}}
+    - \obk{\rho|T_{\ell m}} \obk{\rho|T_{\ell n}}},
+\end{align}
+where we define the inversion operator $\I\equiv\sum_m \p{-1}^m\op{-m}{m}$, and then expand the product $T_{\ell m} T_{\ell n}$ to get
+\begin{align}
+  \C_\ell\sp{\rho} = \sum_{m,n} \I \op{m}{n}
+  \sp{\sum_L g_{\ell mn}^L \obk{\rho|T_{L,m+n}}
+    - \obk{\rho|T_{\ell m}} \obk{\rho|T_{\ell n}}},
+  &&
+  g_{\ell mn}^L \equiv \obk{T_{L,m+n} | T_{\ell m} T_{\ell n}}
+  = f_{\ell m;\ell n}^{L,m+n},
+\end{align}
+where the factors $f_{\ell m;\ell n}^{L,m+n}$ are provided in Appendix \ref{sec:trans_prod}.
 While the error in Eq.~\eqref{eq:error_SM} cannot be known exactly without knowing $\rho$, this error can be estimated by computing the right-hand side of Eq.~\eqref{eq:error_SM} with the replacement $\rho\to\tilde\rho_V$.
@@
 We now seek to maximize the error in Eq.~\eqref{eq:error_SM} over all $\rho$, which will provide an upper bound on reconstruction error.
-To this end, we vary the error with respect to $\rho$ by considering the vector derivative (i.e.~gradient)
+To this end, we vary the error with respect to $\rho$, and consider the vector derivative (i.e.~gradient)
 \begin{align}
@@
   = \f1n \sum_\ell \tr\p{\N_{V\ell}
-    \f{\partial}{\partial\obra\rho} \T_{\rho\ell}}
-  = 0,
+    \f{\partial}{\partial\obra\rho} \C_\ell\sp{\rho}},
   &&
@@
 \begin{align}
-  \f{\partial}{\partial\obra\rho} \T_{\rho\ell}
-  &= \sum_{m,n} \op{m}{n} \otimes \f{\partial}{\partial\obra\rho}
-  \sp{\bk{T_{\ell m}^\dag T_{\ell n}}_\rho
-    - \bk{T_{\ell m}^\dag}_\rho \bk{T_{\ell n}}_\rho} \\
-  &= \sum_{m,n} \op{m}{n} \otimes \f{\partial}{\partial\obra\rho}
-  \sp{\obk{\rho|T_{\ell m}^\dag T_{\ell n}}
-    - \obk{\rho|T_{\ell m}^\dag} \obk{\rho|T_{\ell n}}} \\
-  &= \sum_{m,n} \op{m}{n} \otimes \sp{\oket{T_{\ell m}^\dag T_{\ell n}}
-    - \oket{T_{\ell m}^\dag} \obk{\rho|T_{\ell n}}
-    - \obk{\rho|T_{\ell m}^\dag} \oket{T_{\ell n}}}.
+  \f{\partial}{\partial\obra\rho} \C_\ell\sp{\rho}
+  &= \sum_{m,n} \I \op{m}{n}
+  \f{\partial}{\partial\obra\rho}
+  \sp{\sum_L g_{\ell mn}^L \obk{\rho|T_{L,m+n}}
+    - \obk{\rho|T_{\ell m}} \obk{\rho|T_{\ell n}}} \\
+  &= \sum_{m,n} \I\op{m}{n} \otimes
+  \sp{\sum_L g_{\ell mn}^L \oket{T_{L,m+n}}
+    - \oket{T_{\ell m}} \obk{\rho|T_{\ell n}}
+    - \obk{\rho|T_{\ell m}} \oket{T_{\ell n}}}.
   \label{eq:variation_cov}
@@
 \begin{align}
-  \sum_{m,n} \op{m}{n} \otimes
-  \sp{\oket{T_{\ell m}^\dag} \obk{\rho|T_{\ell n}}
-    + \obk{\rho|T_{\ell m}^\dag} \oket{T_{\ell n}}}
-  &= \sum_{m,n} \op{m}{n} \otimes
-  \sp{\oket{T_{\ell m}^\dag} \obk{T_{\ell n}^\dag|\rho}
-    + \oket{T_{\ell n}} \obk{T_{\ell m}|\rho}} \\
+  \sum_{m,n} \I \op{m}{n} \otimes
+  \sp{\oket{T_{\ell m}} \obk{\rho|T_{\ell n}}
+    + \obk{\rho|T_{\ell m}} \oket{T_{\ell n}}}
+  &= \sum_{m,n} \I \op{m}{n} \otimes
+  \sp{\oket{T_{\ell m}} \obk{T_{\ell n}^\dag|\rho}
+    + \oket{T_{\ell n}} \obk{T_{\ell m}^\dag|\rho}} \\
   &= \sp{\sum_{m,n} \p{\I\op{m}{n}\I + \op{n}{m}}
-    \otimes \oop{T_{\ell m}}{T_{\ell n}}} \oket{\rho},
-\end{align}
-where we define the inversion operator $\I\equiv\sum_m \p{-1}^m\op{-m}{m}$.
-Meanwhile, the constant terms in Eq.~\eqref{eq:variation_cov} are
-\begin{align}
-  \sum_{m,n} \op{m}{n} \otimes \oket{T_{\ell m}^\dag T_{\ell n}}
-  = \sum_{L,m,n} \I \op{m}{n} \otimes g_{\ell mn}^L \oket{T_{L,m+n}},
-  &&
-  g_{\ell mn}^L \equiv \obk{T_{L,m+n}|T_{\ell m} T_{\ell n}}
-  = f_{\ell m;\ell n}^{L,m+n}
-\end{align}
-where the factors $f_{\ell m;\ell n}^{L,m+n}$ are provided in Appendix \ref{sec:trans_prod}.
+    \otimes \oop{T_{\ell m}}{T_{\ell n}}} \oket{\rho}.
+\end{align}
 Altogether, we have that
 \begin{align}
-  \f{\partial}{\partial\obra\rho} \T_{\rho\ell}
+  \f{\partial}{\partial\obra\rho} \C_\ell\sp{\rho}
   = \sum_{m,n} g_{\ell mn}^L \I\op{m}{n} \otimes \oket{T_{L,m+n}}
   - \sp{\sum_{m,n} \p{\op{n}{m} + \I\op{m}{n}\I}
-    \otimes \oop{T_{\ell m}}{T_{\ell n}}} \oket{\rho}.
-\end{align}
-Substituting this result back into Eq.~\eqref{eq:variation}, we find that
+    \otimes \oop{T_{\ell m}}{T_{\ell n}}} \oket{\rho},
+\end{align}
+which we can substitute back into Eq.~\eqref{eq:variation} to find that
 \begin{align}
@@
   \Bbk{\norm{\tilde\rho_V-\rho}^2}
-  = \oket{\chi_V} - \sum_\ell \L_{V\ell} \oket\rho
-  = 0,
+  = \oket{\chi_V} - \sum_\ell \L_{V\ell} \oket\rho,
 \end{align}
@@
 \begin{align}
-  \oket\rho = \sum_\ell \L_{V\ell}^{-1} \oket{\chi_{V\ell}},
+  \oket{\rho^\star} = \sum_\ell \L_{V\ell}^{-1} \oket{\chi_{V\ell}},
 \end{align}
 where $\L_{V\ell}^{-1}$ is the inverse of $\L_{V\ell}$.
-
-[\red{todo: substitute this solution back into the reconstruction error to find an upper bound}]
+In total, the ultimate upper bound on reconstruction error is
+\begin{align}
+  \Bbk{\norm{\tilde\rho_V-\rho}^2} \le
+  \left. \Bbk{\norm{\tilde\rho_V-\rho}^2} \right|_{\rho=\rho^\star}
+  = \f1n \sum_{\ell} \obk{\N_{V\ell}|\C_\ell\sp{\rho^\star}}.
+\end{align}
+
+[\red{todo: finish substituting $\rho^\star$, or otherwise discuss an ``efficient'' calculation of this bound.}]
 

### 2020-11-13 sun_phases/spin_qudit_tomo.tex
added some minor discussions to appendices, renamed the classical / quantum error scales

@@
 \end{align}
-where $\norm{X} \equiv \sqrt{\obk{X|X}} = \sqrt{\tr\p{X^\dag X}}$ is the (Hilbert-Schmidt, or Frobenius) norm of $X$, and
+where $\norm{X} \equiv \sqrt{\obk{X|X}} = \sqrt{\tr\p{X^\dag X}}$ is the (Hilbert-Schmidt, or Frobenius) norm of $X$, and the {\it classical error scale}
 \begin{align}
@@
 is the squared norm of the right inverse $M_V^{-1}$ of $M_V$.
-Note that $\S_V^2$ diverges if $M_V$ is singular, in which case measuring spin projections along all axes in $V$ does not provide sufficient information to reconstruct arbitrary quantum states.
-
-Computing $\S_V^2$ to certify the bound in Eq.~\eqref{eq:bound_eps} requires constructing the measurement matrix $M_V$ and finding its singular values.
+Note that the classical error scale $\S_V^2$ diverges if $M_V$ is singular, in which case measuring spin projections along all axes in $V$ does not provide sufficient information to reconstruct arbitrary quantum states.
+
+Computing the classical error scale $\S_V^2$ to certify the bound in Eq.~\eqref{eq:bound_eps} requires constructing the measurement matrix $M_V$ and finding its singular values.
 The complexity of this task can be greatly reduced by the fact that the degree $\ell$ of a transition operator $T_{\ell m}$ is preserved under rotations, which implies that the unitary
@@
 where $M_{V\ell}^{-1}$ is the right inverse of $M_{V\ell}$.
-Constructing the block $M_{V\ell}$ and computing its singular values takes at most $O(\abs{V}d^2)$ time (see Appendix \ref{sec:compute} for additional details on efficiently computing $\S_{V\ell}^2$ [\red{todo: write tutorial}]).
-Altogether, if we assume that $\abs{V}\sim d$, then computing $\S_V^2$ to certify the bound in Eq.~\eqref{eq:bound_eps} takes $O(d^4)$ serial or $O(d^3)$ parallel runtime (see Figure \ref{fig:times}).
+Constructing the block $M_{V\ell}$ and computing its singular values takes at most $O(\abs{V}d^2)$ time (see Appendix \ref{sec:compute} for details on efficiently computing $\S_{V\ell}^2$).
+If we assume that $\abs{V}\sim d$, then computing $\S_V^2$ to certify the bound in Eq.~\eqref{eq:bound_eps} takes $O(d^4)$ serial or $O(d^3)$ parallel runtime (see Figure \ref{fig:times}).
 
@@
     Dashed line shows a fit to runtime $t=c d^\alpha$ for all $d\ge100$, finding $\alpha\approx 3.7$.
-    These results do not include the $O(d^3)$ runtime for pre-computing $\pi/2$-pulses $e^{\i\p{\pi/2}S_\y}$ (see Appendix \ref{sec:compute}), which can be recycled for every new choice of measurement axes $V$.}
+    These results do not include fixed $O(d^3)$ runtimes for pre-computing quantities that can be recycled for every new choice of measurement axes $V$.}
   \label{fig:times}
@@
 However, this assumption breaks down when measurement error is limited by fundamental quantum (shot) noise.
-We relax the assumption of Eq.~\eqref{eq:error_assumption} in Appendix \ref{sec:error}, where we instead assume that $\tilde\rho_V$ is built from $n$ independent measurements of spin projection along every axis $\v v\in V$.
+We relax the assumption of Eq.~\eqref{eq:error_assumption} in Appendix \ref{sec:error}, where we instead assume that $\tilde\rho_V$ is built from $n$ independent measurements of spin projection along every axis $\v v\in V$, with shot noise the dominant source of error.
 In this case, the mean squared distance between the approximation $\tilde\rho_V$ and the true state $\rho$ can be compactly expressed as as an inner product of the {\it covariance matrices}
@@
 \end{align}
-where $\cov_\rho\p{X,Y} \equiv \bk{XY}_\rho - \bk{X}_\rho \bk{Y}_\rho$, and the {\it shot noise matrices}
+where $\cov_\rho\p{X,Y} \equiv \bk{XY}_\rho - \bk{X}_\rho \bk{Y}_\rho$, and the {\it noise matrices}
 \begin{align}
@@
   \Bbk{\norm{\tilde\rho_V-\rho}^2}
-  = \f1n \sum_\ell \obk{\N_{V\ell}|\C_\ell\sp{\rho}}.
+  = \f1n \sum_{\ell>0} \obk{\N_{V\ell}|\C_\ell\sp{\rho}}.
   \label{eq:error}
@@
 \end{align}
-where the complexity of computing the {\it error scale} $\epsilon_V^2$ is the same as that of computing $\S_V^2$.
+where the complexity of computing the {\it quantum error scale} $\epsilon_V^2$ is the same as that of computing the classical error scale $\S_V^2$.
 For comparison with the bound in Eq.~\eqref{eq:bound_eps}, we note that the error scale $\epsilon_V^2<\S_V^2/2$, so the previous bound still holds with the replacement $\epsilon^2\to1/2n$.
@@
 
-As a final comment, we note that Eq.~\eqref{eq:bound} only provides a pessimistic upper bound on statistical error, which can be calculated without prior knowledge of the true qudit state $\rho$.
+As a final comment, we note that the quantum error scale only provides a pessimistic upper bound on statistical error, which can be calculated without prior knowledge of the true qudit state $\rho$.
 The actual error in the reconstruction $\tilde\rho_V$ of a particular state $\rho$, provided by Eq.~\eqref{eq:error}, may be considerably smaller.
@@
 
-The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and optimize (using any standard minimization algorithm) the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the error scale $\epsilon_V^2$ defined in Eq.~\eqref{eq:scale}.
-If $\abs{V}$ is too large for such optimization  to be practical, one can instead generate many random sets of measurement axes, and then choose the set with the smallest error scale $\epsilon_V^2$.
+The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and optimize (using any standard minimization algorithm) the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the quantum error scale $\epsilon_V^2$ defined in Eq.~\eqref{eq:scale}.
+If $\abs{V}$ is too large for such optimization  to be practical, one can instead generate many random sets of measurement axes, and then choose the set with the smallest quantum error scale $\epsilon_V^2$.
 After choosing a set of measurement axes $V$, (experimentally) make $n$ measurements of spin projection along every axis $\v v\in V$, and set $\tilde\Pi_{\v v\mu}\approx\bk{\Pi_{\v v\mu}}_\rho$ to the fraction of times in which the measurement outcome was $\mu$.
@@
 
-As a last point, we mention that the error scale $\epsilon_V^2$ can also be used to evaluate the ``quality'' of measurement axes $V$ chosen with existing methods for sparse signal recovery on the sphere \cite{rauhut2011sparse, alem2012sparse}.
+As a last point, we mention that the quantum error scale $\epsilon_V^2$ can also be used to evaluate the ``quality'' of measurement axes $V$ chosen with existing methods for sparse signal recovery on the sphere \cite{rauhut2011sparse, alem2012sparse}.
 In practice, we find that uniform random sampling outperforms these methods, in the sense that it generates low-error-scale sets of measurement axes with higher probability.
@@
 Even so, $2d-1$ measurement axes may not be the optimal number to use.
-Increasing the number of measurement axes generally decreases the error scale $\epsilon_V^2$, but at the cost of having to estimate more observables.
+Increasing the number of measurement axes generally decreases the quantum error scale $\epsilon_V^2$, but at the cost of having to estimate more observables.
 At a fixed total number of measurements, increasing the number of measurement axes $\abs{V}$ reduces the number of measurements $n$ devoted to each axis $\v v\in V$.
@@
 \subsection{Exact reconstruction error}
+\label{sec:error_exact}
 
@@
   \Bbk{\norm{\tilde\rho_V-\rho}^2}
-  = \f1n \sum_\ell \obk{\N_{V\ell}|\C_\ell\sp{\rho}}.
+  = \f1n \sum_{\ell>0} \obk{\N_{V\ell}|\C_\ell\sp{\rho}}.
   \label{eq:error_SM}
 \end{align}
+Here we explicitly throw out the $\ell=0$ term because $\C_0\sp{\rho}=0$.
 In practice, to evaluate this error it is useful to expand the covariance matrix as
@@
 where the factors $f_{\ell m;\ell n}^{L,m+n}$ are provided in Appendix \ref{sec:trans_prod}.
-While the error in Eq.~\eqref{eq:error_SM} cannot be known exactly without knowing $\rho$, this error can be estimated by computing the right-hand side of Eq.~\eqref{eq:error_SM} with the replacement $\rho\to\tilde\rho_V$.
+We discuss the details of efficiently computing the error in Eq.~\eqref{eq:error_SM} in Appendix \ref{sec:compute}, but devote the rest of this Appendix to finding upper bounds for this error.
 
@@
 
-We now seek to maximize the error in Eq.~\eqref{eq:error_SM} over all $\rho$, which will provide an upper bound on reconstruction error.
-To this end, we vary the error with respect to $\rho$, and consider the vector derivative (i.e.~gradient)
+We now seek to maximize the error in Eq.~\eqref{eq:error_SM} over all $\rho$, which will provide an upper bound on reconstruction error for a given set of axes $V$.
+To this end, we vary the error with respect to $\rho$, and consider the gradient
 \begin{align}
   \f{\partial}{\partial\obra\rho} \Bbk{\norm{\tilde\rho_V-\rho}^2}
-  = \f1n \sum_\ell \tr\p{\N_{V\ell}
+  = \f1n \sum_{\ell>0} \tr\p{\N_{V\ell}
     \f{\partial}{\partial\obra\rho} \C_\ell\sp{\rho}},
@@
 \end{align}
-Here $\partial/\partial\obra\rho$ is essentially a vector of all $\partial/\partial\rho_{LM}$, which satisfies
+The gradient operator $\partial/\partial\obra\rho$ is essentially a vector of all $\partial/\partial\rho_{LM}$, which satisfies
 \begin{align}
@@
 \end{align}
-The variation in Eq.~\eqref{eq:variation} leads us to consider
+To evaluate Eq.~\eqref{eq:variation}, we first simplify
 \begin{align}
@@
     - \oket{T_{\ell m}} \obk{\rho|T_{\ell n}}
-    - \obk{\rho|T_{\ell m}} \oket{T_{\ell n}}}.
+    - \obk{\rho|T_{\ell m}} \oket{T_{\ell n}}},
   \label{eq:variation_cov}
 \end{align}
+where as a reminder we have defined $\I\equiv\sum_m\p{-1}^m\op{m}$ and $g_{\ell mn}^L\equiv\obk{T_{L,m+n}|T_{\ell m}T_{\ell n}}$.
 The $\sim\rho$ terms in Eq.~\eqref{eq:variation_cov} can be simplified by the fact that
@@
 \end{align}
-Altogether, we have that
+Altogether, we have
 \begin{align}
@@
 \begin{align}
-  \oket{\rho^\star} = \sum_\ell \L_{V\ell}^{-1} \oket{\chi_{V\ell}},
+  \oket{\rho_V^\star} = \sum_\ell \L_{V\ell}^{-1} \oket{\chi_{V\ell}},
 \end{align}
@@
   \Bbk{\norm{\tilde\rho_V-\rho}^2} \le
-  \left. \Bbk{\norm{\tilde\rho_V-\rho}^2} \right|_{\rho=\rho^\star}
-  = \f1n \sum_{\ell} \obk{\N_{V\ell}|\C_\ell\sp{\rho^\star}}.
-\end{align}
-
-[\red{todo: finish substituting $\rho^\star$, or otherwise discuss an ``efficient'' calculation of this bound.}]
+  \left. \Bbk{\norm{\tilde\rho_V-\rho}^2} \right|_{\rho=\rho_V^\star}
+  = \f1n \sum_{\ell>0} \obk{\N_{V\ell}|\C_\ell\sp{\rho_V^\star}}.
+  \label{eq:bound_SM}
+\end{align}
+We discuss the details of efficiently computing this bound in Appendix \ref{sec:compute}, and devote the rest of this Appendix to deriving a simpler (and looser) bound that allows for direct comparison with the naive bound in Eq.~\eqref{eq:bound_eps} of the main text.
+Note that the bound in Eq.~\eqref{eq:bound_SM} is obtained by maximizing reconstruction error over all vectors $\oket\rho$ regardless of their spectrum, which implies that the maximizer $\oket{\rho_V^\star}$ may generally be an unphysical state with negative eigenvalues.
+In principle, there is therefore room for improvement on the bound in Eq.~\eqref{eq:bound_SM} by maximizing reconstruction error over the (restricted) domain of states with a non-negative spectrum.
 
@@
 where $\S_V^2 = \sum_\ell \S_{V\ell}^2$.
+The naive error bound in Eq.~\eqref{eq:bound_eps} of the main text therefore holds in the quantum case with the replacement $\epsilon^2\to1/2n$.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Efficiently computing error bounds}
+\section{Efficiently computing reconstruction errors and bounds}
 \label{sec:compute}
 
-[\red{todo: write a short tutorial on computing the error scale}]
+[\red{todo: write a short tutorial on computing error bounds}]
+
+% We note that when computing this bound in practice, the vector $\oket{\I g_{L\ell}}$ should be pre-computed once and recycled for every new choice of axes $V$, and the inner product $\obk{\N_{V\ell}|\D_M|\I g_{L\ell}}$ to compute $\chi^V_{LM}$ should be seen as an inner product of $O(d)$-sized off-diagonal bands of $\I g_{L\ell}$ and $\N_{V\ell}$.
+% Also, the construction of $\oket{\chi_V}$, $\L_{V\ell}^{-1}$, and $\oket{\rho_V^\star}$ is most efficiently performed in the transition operator basis $\set{\oket{T_{\ell m}}}$.
 

### 2020-11-13 sun_phases/spin_qudit_tomo.tex
added some discussions in sections on reconstruction error, which are now 'complete' (though subject to future revision)

@@
 
+\usepackage{dsfont} % for identity operator
+\newcommand{\1}{\mathds{1}}
+
 %%% figures
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Reconstruction error bound}
+\section{Tomographic reconstruction error}
 \label{sec:bound}
@@
 \end{enumerate*}
-Whereas step \ref{tomo:measure} can involve a host of platform-dependent technical challenges, here we discuss the steps to take before and after collecting measurement data.
+Whereas step \ref{tomo:measure} can involve a host of platform-dependent technical challenges, in the following two sections we discuss the steps to take before and after collecting measurement data.
 
@@
 By a simple change of basis, measuring these projectors is equivalent to measuring the set of transition operators $\v T_V\equiv\set{T_{\v v\ell,0}}$.
-By flattening the operators in $\v T_V$ into column vectors $\oket{T_{\v v\ell,0}}$, we can construct the {\it measurement matrix}
+By flattening the $d\times d$ matrices $T_{\v v\ell,0}$ in $\v T_V$ into $d^2$-component column vectors $\oket{T_{\v v\ell,0}}$, we can construct the {\it measurement matrix}
 \begin{align}
@@
 \end{align}
+Here $\v v$ and $\ell$ label a column of $M_V$, or equivalently label a standard basis vector $\ket{\v v\ell}$ of a $\p{\abs{V}\times d}$-dimensional vector space.
 A necessary and sufficient condition for $V$ to allow for full state tomography is that the measured transition operators in $\v T_V$, or equivalently the columns of $M_V$, span the entire ($d^2$-dimensional) space of operators on a $d$-level spin qudit.
@@
 \begin{align}
-  \rho = \sum_{k=1}^{d^2} \rho_k^V \p{Q_k^V}^\dag,
+  \rho = \sum_{k=1}^{d^2} \rho_k^V Q_k^V,
   &&
@@
 \begin{align}
-  \tilde\rho_V \equiv \sum_k \tilde\rho^V_k \p{Q^V_k}^\dag,
+  \tilde\rho_V \equiv \sum_k \tilde\rho^V_k Q^V_k,
   \label{eq:reconstructed_state}
@@
 \end{align}
-is the squared norm of the right inverse $M_V^{-1}$ of $M_V$.
+is the squared norm of the right inverse $M_V^{-1}$ of $M_V$, such that $M_V M_V^{-1}=1$.
 Note that the classical error scale $\S_V^2$ diverges if $M_V$ is singular, in which case measuring spin projections along all axes in $V$ does not provide sufficient information to reconstruct arbitrary quantum states.
@@
 \begin{align}
-  U \equiv \sum_{\ell,m} \ket{\ell m} \obra{T_{\ell m}},
-\end{align}
-block-diagonalizes the measurement matrix:
-\begin{align}
-  U M_V = \sum_\ell \op{\ell} \otimes M_{V\ell},
+  U \equiv \sum_{\ell=0}^{d-1} \sum_{m=-\ell}^\ell
+  \oket{T_{\ell m}} \bra{\ell m},
+  \label{eq:trans_basis}
+\end{align}
+with vectors $\oket{T_{\ell m}}$ in a column indexed by integers $\p{\ell,m}$, block-diagonalizes the measurement matrix into $d$ blocks indexed by the degree $\ell$:
+\begin{align}
+  U^\dag M_V = \sum_{\ell=0}^{d-1} \op{\ell} \otimes M_{V\ell},
 \end{align}
@@
 
-The assumption that observables can be estimated up to uncorrelated noise with maximal variance $\epsilon^2$, summarized by Eq.~\eqref{eq:error_assumption}, is reasonable when measurement error is dominated by experimental sources of noise.
+The assumption that observables can be estimated up to uncorrelated noise with maximal variance $\epsilon^2$, summarized by Eq.~\eqref{eq:error_assumption}, is reasonable when measurement error is dominated by classical sources of experimental noise.
 However, this assumption breaks down when measurement error is limited by fundamental quantum (shot) noise.
 We relax the assumption of Eq.~\eqref{eq:error_assumption} in Appendix \ref{sec:error}, where we instead assume that $\tilde\rho_V$ is built from $n$ independent measurements of spin projection along every axis $\v v\in V$, with shot noise the dominant source of error.
-In this case, the mean squared distance between the approximation $\tilde\rho_V$ and the true state $\rho$ can be compactly expressed as as an inner product of the {\it covariance matrices}
-\begin{align}
-  \C_\ell\sp{\rho} \equiv \sum_{m,n}
+In this case, the mean squared distance between the approximation $\tilde\rho_V$ and the true state $\rho$ can be compactly expressed using trace inner products of the {\it covariance matrices}
+\begin{align}
+  \C_\ell\sp{\rho} \equiv \sum_{m,n=-\ell}^\ell
   \cov_\rho\p{T_{\ell m}^\dag, T_{\ell n}} \op{m}{n}
@@
 \end{align}
-This reconstruction error can be maximized analytically, yielding a reconstruction error bound of the form
+Maximizing this error over all $\rho$ (for a fixed set of axes $V$) yields a reconstruction error bound of the form
 \begin{align}
@@
 \end{align}
-where the complexity of computing the {\it quantum error scale} $\epsilon_V^2$ is the same as that of computing the classical error scale $\S_V^2$.
-For comparison with the bound in Eq.~\eqref{eq:bound_eps}, we note that the error scale $\epsilon_V^2<\S_V^2/2$, so the previous bound still holds with the replacement $\epsilon^2\to1/2n$.
-
-[\red{todo: finish discussion about $\epsilon_V^2$}]
-
-As a final comment, we note that the quantum error scale only provides a pessimistic upper bound on statistical error, which can be calculated without prior knowledge of the true qudit state $\rho$.
+where the complexity of computing the {\it quantum error scale} $\epsilon_V^2$ (defined below) is the same as that of computing the classical error scale $\S_V^2$ (see Figure \ref{fig:times}).
+For comparison with the bound in Eq.~\eqref{eq:bound_eps}, we note that $\epsilon_V^2<\S_V^2/2$, so the previous bound still holds with the replacement $\epsilon^2\to1/2n$.
+We also note that the (squared) reconstruction error in Eq.~\eqref{eq:error} and its bound in Eq.~\eqref{eq:bound} obey the ``standard quantum limit'' of $\sim1/\sqrt{n}$ scaling in the number of measurements.
+In principle, this scaling can be improved to $\sim1/n$ by preparing and measuring entangled copies of many qudits \cite{giovannetti2006quantum}.
+
+The quantum error scale $\epsilon_V^2$ only provides a pessimistic upper bound on statistical error, which can be calculated without prior knowledge of the true qudit state $\rho$.
 The actual error in the reconstruction $\tilde\rho_V$ of a particular state $\rho$, provided by Eq.~\eqref{eq:error}, may be considerably smaller.
 While the true error in $\tilde\rho_V$ cannot be known exactly without knowing $\rho$, this error can be estimated by computing the right-hand side of Eq.~\eqref{eq:error} with the replacement $\rho\to\tilde\rho_V$.
+
+The remainder of this section is devoted to defining $\epsilon_V^2$, and can be skipped without consequence for the discussions in the rest of this paper.
+To write out the quantum error scale $\epsilon_V^2$ explicitly, we define a matrix of structure constants
+\begin{align}
+  g_{L\ell} \equiv \sum_{m,n} g^L_{\ell mn} \op{m}{n},
+  &&
+  g^L_{\ell mn} \equiv \obk{T_{L,m+n}|T_{\ell m}T_{\ell n}},
+\end{align}
+and the inversion operator $\I\equiv\sum_m\p{-1}^m\op{m}{m}$, which define components of the $d^2$-component vector
+\begin{align}
+  \ket{\chi_V}
+  &\equiv \sum_{L=0}^{d-1} \sum_{M=-L}^L \chi^V_{LM} \ket{LM}, \\
+  \chi^V_{LM} &\equiv \sum_\ell \obk{\N_{V\ell}|\D_M|\I g_{L\ell}}, \\
+  \D_M &\equiv \sum_{m,n} \delta_{M,m+n} \op{mn}.
+\end{align}
+We then define the state vector
+\begin{align}
+  \oket{\rho_V^\star} = \f1d \oket{\1_d} + U \sum_{\ell>0} \sp{\N_{V\ell} + \I\N_{V\ell}^*\I}^{-1} \ket{\chi_{V\ell}},
+\end{align}
+where $\1_d$ is the $d\times d$ identity matrix, $\ket{\chi_{V\ell}} \equiv \sum_{m=-\ell}^\ell \ket{\ell m} \bk{\ell m|\chi_V}$ is the projection of $\ket{\chi_V}$ onto the subspace of fixed degree $\ell$, the unitary $U$ defined in Eq.~\eqref{eq:trans_basis} takes $\ket{\ell m}\to \oket{T_{\ell m}}$, and $\N_{V\ell}^*$ is the complex conjugate of $\N_{V\ell}$.
+The quantum error scale $\epsilon_V^2$ is then
+\begin{align}
+  \epsilon_V^2 \equiv \sum_{\ell>0} \obk{\N_{V\ell}|\C_\ell\sp{\rho_V^\star}}.
+  \label{eq:scale}
+\end{align}
+We derive Eq.~\eqref{eq:scale} in Appendix \ref{sec:error}, and discuss its efficient calculation in Appendix \ref{sec:compute}.
+Note that the ``state'' $\rho_V^\star$, which maximizes the reconstruction error in Eq.~\eqref{eq:error}, has trace $\tr\rho_V^\star=1$, but may generally have negative eigenvalues.
+In principle, there is therefore room for improvement on the bound in Eq.~\eqref{eq:bound} by maximizing reconstruction error over the (restricted) domain of states $\rho$ with a non-negative spectrum.
 
@@
 After choosing a set of measurement axes $V$, (experimentally) make $n$ measurements of spin projection along every axis $\v v\in V$, and set $\tilde\Pi_{\v v\mu}\approx\bk{\Pi_{\v v\mu}}_\rho$ to the fraction of times in which the measurement outcome was $\mu$.
-The estimates $\tilde\Pi_{\v v\mu}$ of $\bk{\Pi_{\v v\mu}}_\rho$ determine estimates of the expectation values $\bk{T_{\v v\ell,0}}_\rho$, which can be used to construct the state $\tilde\rho_V$ defined in Eq.~\eqref{eq:reconstructed_state}.
-If $\tilde\rho_V$ has negative eigenvalues, its (Euclidean) distance from $\rho$ can be reduced with maximum-likelihood corrections \cite{smolin2012efficient}.
+The estimates $\tilde\Pi_{\v v\mu}$ of $\bk{\Pi_{\v v\mu}}_\rho$ determine estimates of the expectation values $\bk{T_{\v v\ell,0}}_\rho$, which can be used to construct an estimate $\tilde\rho_V$ of $\rho$, defined in Eq.~\eqref{eq:reconstructed_state}.
+If $\tilde\rho_V$ has negative eigenvalues, its distance from $\rho$ can be reduced with maximum-likelihood corrections \cite{smolin2012efficient}.
 
@@
 As a last point, we mention that the quantum error scale $\epsilon_V^2$ can also be used to evaluate the ``quality'' of measurement axes $V$ chosen with existing methods for sparse signal recovery on the sphere \cite{rauhut2011sparse, alem2012sparse}.
-In practice, we find that uniform random sampling outperforms these methods, in the sense that it generates low-error-scale sets of measurement axes with higher probability.
+In practice, we find that uniform random sampling outperforms these methods, in the sense that it generates low-error-scale sets of measurement axes with higher probability [\red{todo: add figure}].
 Two observations may explain the poor performance of these sparse signal recovery methods.
@@
 \end{align}
-which is block diagonalized as
-\begin{align}
-  U \equiv \sum_{\ell,m} \ket{\ell m} \obra{T_{\ell m}},
-  &&
-  U M_V = \sum_\ell \op{\ell} \otimes M_{V\ell},
+which can be block diagonalized as
+\begin{align}
+  U^\dag M_V = \sum_\ell \op{\ell} \otimes M_{V\ell},
+  &&
+  U \equiv \sum_{\ell,m} \oket{T_{\ell m}} \bra{\ell m},
   &&
@@
 \begin{align}
-  \rho = \sum_{\ell,m} \bk{Q^V_{\ell m}}_\rho \p{Q^V_{\ell m}}^\dag,
+  \rho = \sum_{\ell,m} \bk{Q^V_{\ell m}}_\rho Q^V_{\ell m},
 \end{align}
@@
 \begin{align}
-  \tilde\rho_V
-  \equiv \sum_{\ell,m} \sp{\sum_i q^V_{\ell mi} \tilde T_{i\ell}}
-  \p{Q^V_{\ell m}}^\dag
-  \approx \sum_{\ell,m} \sp{\sum_i q^V_{\ell mi} \bk{T_{i\ell}}_\rho}
-  \p{Q^V_{\ell m}}^\dag
-  = \sum_{\ell,m} \bk{Q^V_{\ell m}}_\rho \p{Q^V_{\ell m}}^\dag
+  \tilde\rho_V \equiv \sum_{\ell,m}
+  \sp{\sum_i q^V_{\ell mi} \tilde T_{i\ell}} Q^V_{\ell m}
+  \approx \sum_{\ell,m}
+  \sp{\sum_i q^V_{\ell mi} \bk{T_{i\ell}}_\rho} Q^V_{\ell m}
+  = \sum_{\ell,m} \bk{Q^V_{\ell m}}_\rho Q^V_{\ell m}
   = \rho.
@@
 where the factors $f_{\ell m;\ell n}^{L,m+n}$ are provided in Appendix \ref{sec:trans_prod}.
-We discuss the details of efficiently computing the error in Eq.~\eqref{eq:error_SM} in Appendix \ref{sec:compute}, but devote the rest of this Appendix to finding upper bounds for this error.
+We discuss the details of efficiently computing the error in Eq.~\eqref{eq:error_SM} in Appendix \ref{sec:compute}, and devote the rest of this Appendix to finding upper bounds for this error.
 
@@
   \L_{V\ell} \equiv \sum_{m,n}
-  \sp{\bk{m|\N_{V\ell}|n} + \bk{n|\I\N_{V\ell}\I|m}}
+  \bk{m|\p{\N_{V\ell} + \I \N_{V\ell}^* \I}|n}
   \oop{T_{\ell m}}{T_{\ell n}}.
@@
 \end{align}
+Here $\N_{V\ell}^*$ is the complex conjugate (and also the transpose) of $\N_{V\ell}$.
 Defining the projection of $\oket{\chi_V}$ onto the degree-$\ell$ subspace of transition operators,
@@
 \begin{align}
-  \oket{\rho_V^\star} = \sum_\ell \L_{V\ell}^{-1} \oket{\chi_{V\ell}},
-\end{align}
-where $\L_{V\ell}^{-1}$ is the inverse of $\L_{V\ell}$.
+  \oket{\rho_V^\star}
+  = \f1d \oket{\1_d} + \sum_{\ell>0} \L_{V\ell}^{-1} \oket{\chi_{V\ell}},
+\end{align}
+where $\1_d$ is the $d\times d$ identity matrix (added artificially to enforce $\tr\rho_V^\star=1$; the identity component of $\rho_V^\star$ does not actually contribute to reconstruction error), and $\L_{V\ell}^{-1}$ is the inverse of $\L_{V\ell}$.
 In total, the ultimate upper bound on reconstruction error is
@@
 
-In order to establish simpler bounds on reconstruction error, we define the probability $p^i_\mu\equiv\bk{\Pi_{i\mu}}_\rho$, collect these probabilities into the classical probability distribution $p^i = \p{p^i_{-s}, p^i_{-s+1}, \cdots, p^i_s}$, and define the vector $t_\ell \equiv \p{t_{\ell,-s},t_{\ell,-s+1},\cdots,t_{\ell s}}$ of all $t_{\ell\mu}\equiv\bk{\mu|T_{\ell,0}|\mu}$.
+In order to find a simple bound on reconstruction error, we define the probability $p^i_\mu\equiv\bk{\Pi_{i\mu}}_\rho$, collect these probabilities into the classical probability distribution $p^i = \p{p^i_{-s}, p^i_{-s+1}, \cdots, p^i_s}$, and define the vector $t_\ell \equiv \p{t_{\ell,-s},t_{\ell,-s+1},\cdots,t_{\ell s}}$ of all $t_{\ell\mu}\equiv\bk{\mu|T_{\ell,0}|\mu}$.
 We then observe that
@@
   \bbk{\epsilon_{i\ell}^2} = \f1n \times \sigma_{p^i}^2\p{t_\ell},
-\end{align}
-where
-\begin{align}
+  &&
   \sigma_p^2\p{X}
-  \equiv \sum_\mu p_\mu X_\mu^2 - \p{\sum_\mu p_\mu X_\mu}^2
-\end{align}
-is the weighted variance of $X$.
+  \equiv \sum_\mu p_\mu X_\mu^2 - \p{\sum_\mu p_\mu X_\mu}^2,
+\end{align}
+where $\sigma_p^2\p{X}$ is the weighted variance of $X$.
 This variance is maximal when $p$ has equal weight on the largest and smallest values of $X$, which implies that
@@
 \end{align}
-which lets us bound
+and lets us bound
 \begin{align}
@@
   \Gamma_\ell^2 \le \min\Set{\f{2\ell+1}{d}, 1},
-\end{align}
-where $\S_V^2=\sum_\ell\S_{V\ell}^2$.
+  &&
+  \t{and}
+  &&
+  \Bbk{\norm{\tilde\rho_V-\rho}^2} \le \f{\S_V^2}{n}.
+\end{align}
+where $\S_V^2=\sum_\ell\S_{V\ell}^2=\norm{M_V^{-1}}^2$.
 We can get a tighter analytical bound on $\Gamma_\ell^2$ by considering the fact that $t_{\ell\mu}^2 = t_{\ell,-\mu}^2$ due to the symmetries of the Clebsch-Gordan coefficients.
@@
   \Bbk{\norm{\tilde\rho_V-\rho}^2}
-  \le \f1n \sum_\ell \Gamma_\ell^2 \S_{V\ell}^2
-  \le \f{\S_V^2}{2n},
-\end{align}
-where $\S_V^2 = \sum_\ell \S_{V\ell}^2$.
-The naive error bound in Eq.~\eqref{eq:bound_eps} of the main text therefore holds in the quantum case with the replacement $\epsilon^2\to1/2n$.
+  \le \f{\S_V^2}{2n}.
+\end{align}
+The naive error bound in Eq.~\eqref{eq:bound_eps} of the main text therefore still holds with the replacement $\epsilon^2\to1/2n$.
 

### 2020-11-18 sun_phases/spin_qudit_tomo.tex
overhaul of the reconstruction error bound calculation

@@
   \rho = \sum_{\ell=0}^{d-1} \sum_{m=-\ell}^\ell
-  \rho_{\ell m} T_{\ell m}^\dag,
-  &&
-  \rho_{\ell m} \equiv \bk{T_{\ell m}}_\rho,
+  \rho_{\ell m} T_{\ell m},
+  &&
+  \rho_{\ell m} \equiv \bk{T_{\ell m}^\dag}_\rho,
   \label{eq:trans_state}
 \end{align}
-where $\bk{X}_\rho\equiv\obk{\rho|X}=\tr\p{\rho X}$, and $\rho^\dag=\rho$ implies that $\rho_{\ell m}^*=\p{-1}^m\rho_{\ell,-m}$.
+where $\bk{X}_\rho\equiv\tr\p{\rho X}$, and $\rho^\dag=\rho$ implies that $\rho_{\ell m}^*=\p{-1}^m\rho_{\ell,-m}$.
 The transition operators can be physically interpreted as the transition $\ket\psi\to T_{\ell m}\ket\psi$ induced on the state $\ket\psi$ of a spin-$s$ system by the absorption of a spin-$\ell$ particle with spin projection $m$ onto a fixed quantization axis.
@@
 
-Reconstructing an arbitrary density operator $\rho$ essentially requires us to find a set of axes $V=\set{\v v}$ and associated coefficients $C_{\ell mk}\p{\v v}$ that would allow us to recover any coefficient $\rho_{\ell m}$ in expansion in Eq.~\eqref{eq:trans_state} through
-\begin{align}
-  \rho_{\ell m} = \bk{T_{\ell m}}_\rho
-  = \sum_{\v v,k} C_{\ell mk}\p{\v v} \bk{T_{\v vk,0}}_\rho.
+Reconstructing an arbitrary density operator $\rho$ essentially requires us to find a set of axes $V=\set{\v v}$ and associated coefficients $C_{\ell mk}\p{\v v}$ that would allow us to recover any coefficient $\rho_{\ell m}$ in the expansion in Eq.~\eqref{eq:trans_state} through
+\begin{align}
+  \rho_{\ell m} = \bk{T_{\ell m}^\dag}_\rho
+  = \sum_{\v v,k} C_{\ell mk}\p{\v v} \bk{T_{\v vk,0}^\dag}_\rho.
   \label{eq:state_recon}
@@
 \begin{align}
-  T_{\ell m}
-  = \sum_{\v v,k,n} C_{\ell mk}\p{\v v} D^k_{n,0}\p{\v v} T_{kn}.
+  T_{\ell m}^\dag
+  = \sum_{\v v,k,n} C_{\ell mk}\p{\v v}
+  D^k_{n,0}\p{\v v}^* T_{kn}^\dag.
 \end{align}
@@
 \begin{align}
-  \sum_{\v v} C_{\ell m}\p{\v v} D^\ell_{n,0}\p{\v v} = \delta_{mn}
+  \sum_{\v v} C_{\ell m}\p{\v v} D^\ell_{n,0}\p{\v v}^* = \delta_{mn}
   \label{eq:tomo_recovery}
@@
 We will refer to Eq.~\eqref{eq:full_recovery} as the {\it full recovery problem}, and Eq.~\eqref{eq:tomo_recovery} as the the {\it reduced recovery problem}.
-Due to the fact that $D^k_{n,0}\p{\v v} = \sqrt{\frac{4\pi}{2\ell+1}}\, Y_{kn}\p{\v v}^*$, any solution to the full recovery problem automatically solves the reduced recovery problem by setting $C_{\ell m}\p{\v v} = \sqrt{\frac{2\ell+1}{4\pi}}\, \tilde C_{\ell m}\p{\v v}^*$.
+Due to the fact that $D^k_{n,0}\p{\v v}^* = \sqrt{\frac{4\pi}{2\ell+1}}\, Y_{kn}\p{\v v}$, any solution to the full recovery problem automatically solves the reduced recovery problem by setting $C_{\ell m}\p{\v v} = \sqrt{\frac{2\ell+1}{4\pi}}\, \tilde C_{\ell m}\p{\v v}$.
 In principle, this mapping allows us to import a host of existing signal recovery algorithms \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality} for the task of spin qudit tomography.
@@
 one might expect that $2d-1$ suitably chosen measurement axes would always suffice to perform full tomography of a $d$-level spin qudit.
-In the language of classical signal recovery, we would expect there to exist a fixed set of $2\ell_{\t{max}}+1$ points on the sphere that enable the recovery of {\it any} function with fixed degree $\ell\le\ell_{\t{max}}$ (given prior knowledge of $\ell$) \footnote{We could further conjecture that if {\it any} set of points $V\subset\Omega_2$ enables the recovery of degree-$\ell_{\t{max}}$ functions on the sphere, then $V$ would allow recovery degree-$\ell$ functions with $\ell\le\ell_{\t{max}}$.
-However, this condition is stronger than that necessary to guarantee full recovery of a $d$-level spin qudit state with $2d-1$ measurement axes.}.
+In the language of classical signal recovery, we would expect there to exist a fixed set of $2\ell_{\t{max}}+1$ points on the sphere that enable the recovery of {\it any} function with fixed degree $\ell\le\ell_{\t{max}}$ (given prior knowledge of $\ell$) \footnote{We could further conjecture that if {\it any} set of points $V\subset\Omega_2$ enables the recovery of degree-$\ell_{\t{max}}$ functions on the sphere, then $V$ enables the recovery degree-$\ell$ functions with $\ell\le\ell_{\t{max}}$.
+However, this condition is stronger than necessary to guarantee full recovery of a $d$-level spin qudit state with $2d-1$ measurement axes, which only requires the existence of {\it some} such set of points.}.
 Unfortunately, we have no general proof of this conjecture, which requires a deeper analysis of fixed-degree signal recovery on the sphere.
@@
 where $M_{V\ell}^{-1}$ is the right inverse of $M_{V\ell}$.
-Constructing the block $M_{V\ell}$ and computing its singular values takes at most $O(\abs{V}d^2)$ time (see Appendix \ref{sec:compute} for details on efficiently computing $\S_{V\ell}^2$).
+Constructing the block $M_{V\ell}$ and computing its singular values takes at most $O(\abs{V}d^2)$ time.
 If we assume that $\abs{V}\sim d$, then computing $\S_V^2$ to certify the bound in Eq.~\eqref{eq:bound_eps} takes $O(d^4)$ serial or $O(d^3)$ parallel runtime (see Figure \ref{fig:times}).
@@
   \includegraphics{error_scale/error_times.pdf}
-  \caption{Serial runtime to compute $\S_V$ with $\abs{V}=2d-1$ randomly chosen measurement axes, averaged over 100 calculations of $\S_V$ or 5 minutes of runtime (for each $d$), whichever comes first.
-    Dashed line shows a fit to runtime $t=c d^\alpha$ for all $d\ge100$, finding $\alpha\approx 3.7$.
+  \caption{Serial runtime to compute the error scales $\S_V^2$ and $\epsilon_V^2$ with $\abs{V}=2d-1$ randomly chosen measurement axes.
+    Each point is an average over 100 calculations or 5 minutes of runtime, whichever comes first.
+    Dashed lines shows a fit to runtime $t=c d^\alpha$ for all $d\ge100$, finding $\alpha\approx 3.7$.
     These results do not include fixed $O(d^3)$ runtimes for pre-computing quantities that can be recycled for every new choice of measurement axes $V$.}
@@
   \Bbk{\norm{\tilde\rho_V-\rho}^2}
-  = \f1n \sum_{\ell>0} \obk{\N_{V\ell}|\C_\ell\sp{\rho}}.
+  = \f1n \sum_\ell \obk{\N_{V\ell}|\C_\ell\sp{\rho}}.
   \label{eq:error}
@@
 
-The remainder of this section is devoted to defining $\epsilon_V^2$, and can be skipped without consequence for the discussions in the rest of this paper.
+The remainder of this section is devoted to defining $\epsilon_V^2$, which is derived in Appendix \ref{sec:error}, and can be skipped without consequence for the discussions in the rest of this paper.
 To write out the quantum error scale $\epsilon_V^2$ explicitly, we define a matrix of structure constants
 \begin{align}
-  g_{L\ell} \equiv \sum_{m,n} g^L_{\ell mn} \op{m}{n},
-  &&
-  g^L_{\ell mn} \equiv \obk{T_{L,m+n}|T_{\ell m}T_{\ell n}},
-\end{align}
-and the inversion operator $\I\equiv\sum_m\p{-1}^m\op{m}{m}$, which define components of the $d^2$-component vector
-\begin{align}
-  \ket{\chi_V}
-  &\equiv \sum_{L=0}^{d-1} \sum_{M=-L}^L \chi^V_{LM} \ket{LM}, \\
-  \chi^V_{LM} &\equiv \sum_\ell \obk{\N_{V\ell}|\D_M|\I g_{L\ell}}, \\
-  \D_M &\equiv \sum_{m,n} \delta_{M,m+n} \op{mn}.
-\end{align}
-We then define the state vector
-\begin{align}
-  \oket{\rho_V^\star} = \f1d \oket{\1_d} + U \sum_{\ell>0} \sp{\N_{V\ell} + \I\N_{V\ell}^*\I}^{-1} \ket{\chi_{V\ell}},
-\end{align}
-where $\1_d$ is the $d\times d$ identity matrix, $\ket{\chi_{V\ell}} \equiv \sum_{m=-\ell}^\ell \ket{\ell m} \bk{\ell m|\chi_V}$ is the projection of $\ket{\chi_V}$ onto the subspace of fixed degree $\ell$, the unitary $U$ defined in Eq.~\eqref{eq:trans_basis} takes $\ket{\ell m}\to \oket{T_{\ell m}}$, and $\N_{V\ell}^*$ is the complex conjugate of $\N_{V\ell}$.
-The quantum error scale $\epsilon_V^2$ is then
-\begin{align}
-  \epsilon_V^2 \equiv \sum_{\ell>0} \obk{\N_{V\ell}|\C_\ell\sp{\rho_V^\star}}.
-  \label{eq:scale}
-\end{align}
-We derive Eq.~\eqref{eq:scale} in Appendix \ref{sec:error}, and discuss its efficient calculation in Appendix \ref{sec:compute}.
-Note that the ``state'' $\rho_V^\star$, which maximizes the reconstruction error in Eq.~\eqref{eq:error}, has trace $\tr\rho_V^\star=1$, but may generally have negative eigenvalues.
-In principle, there is therefore room for improvement on the bound in Eq.~\eqref{eq:bound} by maximizing reconstruction error over the (restricted) domain of states $\rho$ with a non-negative spectrum.
+  g_{L\ell} \equiv \sum_{m,n=-\ell}^\ell
+  \obk{T_{L,m+n}|T_{\ell m}T_{\ell n}} \op{m}{n},
+\end{align}
+and the inversion operator $\I\equiv\sum_m\p{-1}^m\op{-m}{m}$, which are used to define the $\p{2\ell+1}$-component vectors
+\begin{align}
+  \ket{\chi_{VL}}
+  &\equiv \sum_{M=-L}^L \chi^V_{LM} \ket{M}, \\
+  \chi^V_{Lm} &\equiv \sum_\ell \obk{\N_{V\ell}|\D_M|\I g_{L\ell}}, \\
+  \D_M &\equiv \sum_{m,m'} \delta_{M,m'-m} \op{mm'}.
+\end{align}
+Here $\D_M$ simply picks off the $M$-th diagonal of the matrix it acts on, such that $\obk{\N_{V\ell}|\D_M|\I g_{L\ell}}$ is an inner product between the $M$-th diagonals of $\N_{V\ell}$ and $\I g_{L\ell}$.
+We then define
+\begin{align}
+  \L_{V\ell} \equiv \N_{V\ell} + \I\N_{V\ell}\I,
+\end{align}
+in terms of which we have that $\epsilon_V^2 \equiv \sum_\ell \epsilon_{V\ell}^2$, with
+\begin{align}
+  \epsilon_{V\ell}^2 \equiv
+  \bk{\chi_{V\ell}|\p{\L_{V\ell}^{-1}
+      - \L_{V\ell}^{-1}\N_{V\ell}\L_{V\ell}^{-1}}|\chi_{V\ell}}.
+\end{align}
 
@@
 
-The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and optimize (using any standard minimization algorithm) the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the quantum error scale $\epsilon_V^2$ defined in Eq.~\eqref{eq:scale}.
+The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and optimize (using any standard minimization algorithm) the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the quantum error scale $\epsilon_V^2$.
 If $\abs{V}$ is too large for such optimization  to be practical, one can instead generate many random sets of measurement axes, and then choose the set with the smallest quantum error scale $\epsilon_V^2$.
 After choosing a set of measurement axes $V$, (experimentally) make $n$ measurements of spin projection along every axis $\v v\in V$, and set $\tilde\Pi_{\v v\mu}\approx\bk{\Pi_{\v v\mu}}_\rho$ to the fraction of times in which the measurement outcome was $\mu$.
-The estimates $\tilde\Pi_{\v v\mu}$ of $\bk{\Pi_{\v v\mu}}_\rho$ determine estimates of the expectation values $\bk{T_{\v v\ell,0}}_\rho$, which can be used to construct an estimate $\tilde\rho_V$ of $\rho$, defined in Eq.~\eqref{eq:reconstructed_state}.
+The estimates $\tilde\Pi_{\v v\mu}$ of $\bk{\Pi_{\v v\mu}}_\rho$ determine estimates of the expectation values $\bk{T_{\v v\ell,0}}_\rho$, which can be used to construct an estimate $\tilde\rho_V$ of $\rho$.
 If $\tilde\rho_V$ has negative eigenvalues, its distance from $\rho$ can be reduced with maximum-likelihood corrections \cite{smolin2012efficient}.
@@
   = \cov_\rho\p{T_{i\ell}^\dag,T_{i\ell}}
-  = \sum_{m,n} D^\ell_{m,0}\p{\v v_i}^* D^\ell_{n,0}\p{\v v_i}
-  \cov_\rho\p{T_{\ell m}^\dag, T_{\ell n}},
+  = \sum_{m,m'} D^\ell_{m,0}\p{\v v_i}^* D^\ell_{m',0}\p{\v v_i}
+  \cov_\rho\p{T_{\ell m}^\dag, T_{\ell m'}},
 \end{align}
@@
   \Bbk{\norm{\tilde\rho_V-\rho}^2}
-  = \f1n \sum_{i,\ell,m,n}
+  = \f1n \sum_{i,\ell,m,m'}
   \abs{\tilde{\v q}_{\ell i}}^2
-  D^\ell_{m,0}\p{\v v_i}^* D^\ell_{n,0}\p{\v v_i}
-  \cov_\rho\p{T_{\ell m}^\dag, T_{\ell n}}.
+  D^\ell_{m,0}\p{\v v_i}^* D^\ell_{m',0}\p{\v v_i}
+  \cov_\rho\p{T_{\ell m}^\dag, T_{\ell m'}}.
 \end{align}
@@
   \C_\ell\sp{\rho} \equiv
-  \sum_{m,n} \cov_\rho\p{T_{\ell m}^\dag, T_{\ell n}} \op{m}{n},
+  \sum_{m,m'} \cov_\rho\p{T_{\ell m}^\dag, T_{\ell m'}} \op{m}{m'},
 \end{align}
@@
   \Bbk{\norm{\tilde\rho_V-\rho}^2}
-  = \f1n \sum_{\ell>0} \obk{\N_{V\ell}|\C_\ell\sp{\rho}}.
+  = \f1n \sum_\ell \obk{\N_{V\ell}|\C_\ell\sp{\rho}}.
   \label{eq:error_SM}
 \end{align}
-Here we explicitly throw out the $\ell=0$ term because $\C_0\sp{\rho}=0$.
-In practice, to evaluate this error it is useful to expand the covariance matrix as
-\begin{align}
-  \C_\ell\sp{\rho}
-  = \sum_{m,n} \op{m}{n} \sp{\obk{\rho|T_{\ell m}^\dag T_{\ell n}}
-    - \obk{\rho|T_{\ell m}^\dag} \obk{\rho|T_{\ell n}}}
-  = \sum_{m,n} \I\op{m}{n} \sp{\obk{\rho|T_{\ell m} T_{\ell n}}
-    - \obk{\rho|T_{\ell m}} \obk{\rho|T_{\ell n}}},
-\end{align}
-where we define the inversion operator $\I\equiv\sum_m \p{-1}^m\op{-m}{m}$, and then expand the product $T_{\ell m} T_{\ell n}$ to get
-\begin{align}
-  \C_\ell\sp{\rho} = \sum_{m,n} \I \op{m}{n}
-  \sp{\sum_L g_{\ell mn}^L \obk{\rho|T_{L,m+n}}
-    - \obk{\rho|T_{\ell m}} \obk{\rho|T_{\ell n}}},
-  &&
-  g_{\ell mn}^L \equiv \obk{T_{L,m+n} | T_{\ell m} T_{\ell n}}
-  = f_{\ell m;\ell n}^{L,m+n},
-\end{align}
-where the factors $f_{\ell m;\ell n}^{L,m+n}$ are provided in Appendix \ref{sec:trans_prod}.
-We discuss the details of efficiently computing the error in Eq.~\eqref{eq:error_SM} in Appendix \ref{sec:compute}, and devote the rest of this Appendix to finding upper bounds for this error.
 
@@
 We now seek to maximize the error in Eq.~\eqref{eq:error_SM} over all $\rho$, which will provide an upper bound on reconstruction error for a given set of axes $V$.
-To this end, we vary the error with respect to $\rho$, and consider the gradient
-\begin{align}
-  \f{\partial}{\partial\obra\rho} \Bbk{\norm{\tilde\rho_V-\rho}^2}
-  = \f1n \sum_{\ell>0} \tr\p{\N_{V\ell}
-    \f{\partial}{\partial\obra\rho} \C_\ell\sp{\rho}},
-  &&
-  \f{\partial}{\partial\obra\rho}
-  \equiv \sum_{L,M} \oket{T_{LM}} \f{\partial}{\partial\rho_{LM}}.
-  \label{eq:variation}
-\end{align}
-The gradient operator $\partial/\partial\obra\rho$ is essentially a vector of all $\partial/\partial\rho_{LM}$, which satisfies
-\begin{align}
-  \f{\partial}{\partial\obra\rho} \obk{\rho|X}
-  = \f{\partial}{\partial\obra\rho}
-  \sum_{\ell,m} \obk{\rho|T_{\ell m}} \obk{T_{\ell m}|X}
-  = \sum_{L,M} \oket{T_{LM}} \f{\partial}{\partial\rho_{LM}}
-  \sum_{\ell,m} \rho_{\ell m} \obk{T_{\ell m}|X}
-  = \sum_{L,M} \oket{T_{LM}} \obk{T_{LM}|X}
-  = \oket{X}.
-\end{align}
-To evaluate Eq.~\eqref{eq:variation}, we first simplify
-\begin{align}
-  \f{\partial}{\partial\obra\rho} \C_\ell\sp{\rho}
-  &= \sum_{m,n} \I \op{m}{n}
-  \f{\partial}{\partial\obra\rho}
-  \sp{\sum_L g_{\ell mn}^L \obk{\rho|T_{L,m+n}}
-    - \obk{\rho|T_{\ell m}} \obk{\rho|T_{\ell n}}} \\
-  &= \sum_{m,n} \I\op{m}{n} \otimes
-  \sp{\sum_L g_{\ell mn}^L \oket{T_{L,m+n}}
-    - \oket{T_{\ell m}} \obk{\rho|T_{\ell n}}
-    - \obk{\rho|T_{\ell m}} \oket{T_{\ell n}}},
-  \label{eq:variation_cov}
-\end{align}
-where as a reminder we have defined $\I\equiv\sum_m\p{-1}^m\op{m}$ and $g_{\ell mn}^L\equiv\obk{T_{L,m+n}|T_{\ell m}T_{\ell n}}$.
-The $\sim\rho$ terms in Eq.~\eqref{eq:variation_cov} can be simplified by the fact that
-\begin{align}
-  \obk{\rho|T_{\ell m}}
-  = \tr\p{\rho T_{\ell m}}
-  = \p{-1}^m \tr\p{T_{\ell,-m}^\dag \rho}
-  = \p{-1}^m \obk{T_{\ell,-m}|\rho}
-  = \obk{T_{\ell m}^\dag|\rho},
+To this end, we will first rewrite this error in a way that is more convenient to evaluate for any given $\rho$.
+Using the fact that
+\begin{align}
+  \bk{T_{\ell m}^\dag}_\rho
+  = \obk{\rho|T_{\ell m}^\dag}
+  = \tr\p{\rho T_{\ell m}^\dag}
+  = \tr\p{T_{\ell m}^\dag \rho}
+  = \obk{T_{\ell m}|\rho},
+\end{align}
+we can expand the covariance matrix as
+\begin{align}
+  \C_\ell\sp{\rho}
+  &= \sum_{m,m'} \op{m}{m'} \sp{\obk{\rho|T_{\ell m}^\dag T_{\ell m'}}
+    - \obk{\rho|T_{\ell m}^\dag} \obk{\rho|T_{\ell m'}}} \\
+  &= \sum_{m,m'} \op{m}{m'} \sp{\obk{T_{\ell m'}^\dag T_{\ell m}|\rho}
+    - \obk{T_{\ell m}|\rho} \obk{T_{\ell m'}^\dag|\rho}} \\
+  &= \sum_{m,m'} \op{m}{m'}\I \sp{\obk{T_{\ell m'} T_{\ell m}|\rho}
+    - \obk{T_{\ell m}|\rho} \obk{T_{\ell m'}|\rho}}
+\end{align}
+where we define the inversion operator $\I\equiv\sum_m \p{-1}^m\op{-m}{m}$.
+We can then expand the product $T_{\ell m} T_{\ell m'}$ to simplify
+\begin{align}
+  \obk{T_{\ell m'} T_{\ell m}|\rho}
+  = \sum_L g_{\ell m'm}^L \obk{T_{L,m'+m}|\rho},
+  &&
+  g_{\ell m'm}^L \equiv \obk{T_{L,m'+m} | T_{\ell m'} T_{\ell m}}
+  = f_{\ell m';\ell m}^{L,m'+m},
+\end{align}
+where the (real) factors $f_{\ell m';\ell m}^{L,m'+m}$ are provided in Appendix \ref{sec:trans_prod}.
+Substituting the covariance matrix back into Eq.~\eqref{eq:error_SM} and replacing $\obk{T_{\ell m}|\rho}\to\rho_{\ell m}$, we get
+\begin{align}
+  n \Bbk{\norm{\tilde\rho_V-\rho}^2}
+  = \sum_{L,M} \p{\chi^V_{LM}}^* \rho_{LM}
+  - \sum_{\ell,m,m'} \bk{m'|\I\N_{V\ell}|m}
+  \rho_{\ell m'} \rho_{\ell m},
+  \label{eq:error_comps}
+\end{align}
+where
+\begin{align}
+  \chi^V_{LM}
+  &\equiv \sum_{\ell,m,m'} \delta_{M,m'+m}
+  \bk{m'|\I\N_{V\ell}|m}^* g^L_{\ell m'm} \\
+  &= \sum_{\ell,m,m'} \delta_{M,-m'+m}
+  \bk{m|\N_{V\ell}|m'} \p{-1}^{m'} g^L_{\ell,-m',m} \\
+  &= \sum_\ell \obk{\N_{V\ell}|\D_M|\I g_{L\ell}}
+\end{align}
+is expressed in terms of the matrices
+\begin{align}
+  g_{L\ell} \equiv \sum_{m,m'} g^L_{\ell m'm} \op{m'}{m},
+  &&
+  \D_M \equiv \sum_{m,m'} \delta_{M,-m'+m} \op{m'm}.
+\end{align}
+Here $\D_M$ simply picks off the $M$-th diagonal of the matrix it acts on.
+In total, reconstruction error can be written in the form
+\begin{align}
+  n \Bbk{\norm{\tilde\rho_V-\rho}^2}
+  = \obk{\chi_V|\rho}
+  - \sum_\ell \obk{\rho|U_\ell\N_{V\ell}U_\ell^\dag|\rho},
+  &&
+  \oket{\chi_V} \equiv \sum_{L,M} \chi^V_{LM} \oket{T_{LM}},
+  &&
+  U_\ell \equiv \sum_m \oket{T_{\ell m}} \bra{m}.
+\end{align}
+Performing the change of basis $\oket{T_{\ell m}}\to\ket{\ell m}$ by defining
+\begin{align}
+  \ket{\rho} \equiv U^\dag \oket{\rho}
+  = \sum_{\ell,m} \rho_{\ell m}\ket{\ell m},
+  &&
+  \ket{\chi_V} \equiv U^\dag \oket{\chi_V}
+  = \sum_{\ell,m} \chi^V_{\ell m} \ket{\ell m},
+  &&
+  U \equiv \sum_{\ell,m} \oket{T_{\ell m}} \bra{\ell m},
+\end{align}
+as well as the degree-$\ell$ projections
+\begin{align}
+  \ket{\rho_\ell} \equiv \sum_m \rho_{\ell m} \ket{m},
+  &&
+  \ket{\chi_{V\ell}} \equiv \sum_m \chi^V_{\ell m} \ket{m},
+\end{align}
+we can more simply write
+\begin{align}
+  \Bbk{\norm{\tilde\rho_V-\rho}^2}
+  = \f1n \sum_\ell \sp{\bk{\chi_{V\ell}|\rho_\ell}
+    - \bk{\rho_\ell|\N_{V\ell}|\rho_\ell}}.
+\end{align}
+Given the expansion in Eq.~\eqref{eq:error_comps}, this error is straightforward to maximize by considering its variation with respect to $\rho_{\ell m}$:
+\begin{align}
+  n\, \f{\partial}{\partial\rho_{\ell m}}
+  \Bbk{\norm{\tilde\rho_V-\rho}^2}
+  = \p{\chi^V_{\ell m}}^*
+  - \sum_{m'} \sp{\bk{m'|\I\N_{V\ell}|m} + \bk{m'|\I\N_{V\ell}|m}}
+  \rho_{\ell m'}.
+\end{align}
+The complex conjugate of this variation implies that the state $\rho_V^\star$ that maximizes reconstruction error satisfies
+\begin{align}
+  \ket{\chi_{V\ell}} - \L_{V\ell} \ket{\rho_{V\ell}^\star} = 0,
+  &&
+  \L_{V\ell} \equiv \N_{V\ell} + \I N_{V\ell} \I,
 \end{align}
@@
 \begin{align}
-  \sum_{m,n} \I \op{m}{n} \otimes
-  \sp{\oket{T_{\ell m}} \obk{\rho|T_{\ell n}}
-    + \obk{\rho|T_{\ell m}} \oket{T_{\ell n}}}
-  &= \sum_{m,n} \I \op{m}{n} \otimes
-  \sp{\oket{T_{\ell m}} \obk{T_{\ell n}^\dag|\rho}
-    + \oket{T_{\ell n}} \obk{T_{\ell m}^\dag|\rho}} \\
-  &= \sp{\sum_{m,n} \p{\I\op{m}{n}\I + \op{n}{m}}
-    \otimes \oop{T_{\ell m}}{T_{\ell n}}} \oket{\rho}.
-\end{align}
-Altogether, we have
-\begin{align}
-  \f{\partial}{\partial\obra\rho} \C_\ell\sp{\rho}
-  = \sum_{m,n} g_{\ell mn}^L \I\op{m}{n} \otimes \oket{T_{L,m+n}}
-  - \sp{\sum_{m,n} \p{\op{n}{m} + \I\op{m}{n}\I}
-    \otimes \oop{T_{\ell m}}{T_{\ell n}}} \oket{\rho},
-\end{align}
-which we can substitute back into Eq.~\eqref{eq:variation} to find that
-\begin{align}
-  n\, \f{\partial}{\partial\obra\rho}
+  \ket{\rho_{V\ell}^\star} = \L_{V\ell}^{-1} \ket{\chi_{V\ell}},
+\end{align}
+and in turn
+\begin{align}
   \Bbk{\norm{\tilde\rho_V-\rho}^2}
-  = \oket{\chi_V} - \sum_\ell \L_{V\ell} \oket\rho,
+  \le \Bbk{\norm{\tilde\rho_V-\rho_V^\star}^2}
+  = \f{\epsilon_V^2}{n},
+  \label{eq:bound_SM}
 \end{align}
@@
 \begin{align}
-  \oket{\chi_V} \equiv \sum_{L,M} \chi^V_{LM} \oket{T_{LM}},
-  &&
-  \L_{V\ell} \equiv \sum_{m,n}
-  \bk{m|\p{\N_{V\ell} + \I \N_{V\ell}^* \I}|n}
-  \oop{T_{\ell m}}{T_{\ell n}}.
-\end{align}
-\begin{align}
-  \chi^V_{LM} \equiv \sum_\ell \obk{\N_{V\ell}|\D_M|\I g_{L\ell}},
-  &&
-  g_{L\ell} \equiv \sum_{m,n} g^L_{\ell mn} \op{m}{n},
-  &&
-  \D_M \equiv \sum_{m,n} \delta_{M,m+n} \op{mn}{mn}.
-\end{align}
-Here $\N_{V\ell}^*$ is the complex conjugate (and also the transpose) of $\N_{V\ell}$.
-Defining the projection of $\oket{\chi_V}$ onto the degree-$\ell$ subspace of transition operators,
-\begin{align}
-  \oket{\chi_{V\ell}}
-  \equiv \sum_m \oket{T_{\ell m}} \obk{T_{\ell m}|\chi_V},
-\end{align}
-we thus find that reconstruction error is maximized by the state
-\begin{align}
-  \oket{\rho_V^\star}
-  = \f1d \oket{\1_d} + \sum_{\ell>0} \L_{V\ell}^{-1} \oket{\chi_{V\ell}},
-\end{align}
-where $\1_d$ is the $d\times d$ identity matrix (added artificially to enforce $\tr\rho_V^\star=1$; the identity component of $\rho_V^\star$ does not actually contribute to reconstruction error), and $\L_{V\ell}^{-1}$ is the inverse of $\L_{V\ell}$.
-In total, the ultimate upper bound on reconstruction error is
-\begin{align}
-  \Bbk{\norm{\tilde\rho_V-\rho}^2} \le
-  \left. \Bbk{\norm{\tilde\rho_V-\rho}^2} \right|_{\rho=\rho_V^\star}
-  = \f1n \sum_{\ell>0} \obk{\N_{V\ell}|\C_\ell\sp{\rho_V^\star}}.
-  \label{eq:bound_SM}
-\end{align}
-We discuss the details of efficiently computing this bound in Appendix \ref{sec:compute}, and devote the rest of this Appendix to deriving a simpler (and looser) bound that allows for direct comparison with the naive bound in Eq.~\eqref{eq:bound_eps} of the main text.
-Note that the bound in Eq.~\eqref{eq:bound_SM} is obtained by maximizing reconstruction error over all vectors $\oket\rho$ regardless of their spectrum, which implies that the maximizer $\oket{\rho_V^\star}$ may generally be an unphysical state with negative eigenvalues.
+  \epsilon_V^2 \equiv \sum_\ell \epsilon_{V\ell}^2,
+  &&
+  \epsilon_{V\ell}^2 \equiv \bk{\chi_{V\ell}|\p{\L_{V\ell}^{-1}
+      - \L_{V\ell}^{-1}\N_{V\ell}\L_{V\ell}^{-1}}|\chi_{V\ell}}.
+\end{align}
+Note that the bound in Eq.~\eqref{eq:bound_SM} is obtained by maximizing reconstruction error over all possible $\rho$ regardless of their spectrum, which implies that the maximizer $\rho_V^\star$ may generally be an unphysical state with negative eigenvalues.
 In principle, there is therefore room for improvement on the bound in Eq.~\eqref{eq:bound_SM} by maximizing reconstruction error over the (restricted) domain of states with a non-negative spectrum.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Efficiently computing reconstruction errors and bounds}
-\label{sec:compute}
-
-[\red{todo: write a short tutorial on computing error bounds}]
-
-% We note that when computing this bound in practice, the vector $\oket{\I g_{L\ell}}$ should be pre-computed once and recycled for every new choice of axes $V$, and the inner product $\obk{\N_{V\ell}|\D_M|\I g_{L\ell}}$ to compute $\chi^V_{LM}$ should be seen as an inner product of $O(d)$-sized off-diagonal bands of $\I g_{L\ell}$ and $\N_{V\ell}$.
-% Also, the construction of $\oket{\chi_V}$, $\L_{V\ell}^{-1}$, and $\oket{\rho_V^\star}$ is most efficiently performed in the transition operator basis $\set{\oket{T_{\ell m}}}$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Transition operator product expansion}
@@
 \end{align}
-we can use the fact that $2\ell_2$ is (in our case) always even to expand
+we can use the fact that $2\ell_2$ is always even (because $\ell_2$ is always an integer) to expand
 \begin{align}

### 2020-11-18 sun_phases/spin_qudit_tomo.tex
minor cleanup and notation changes

@@
 
-The remainder of this section is devoted to defining $\epsilon_V^2$, which is derived in Appendix \ref{sec:error}, and can be skipped without consequence for the discussions in the rest of this paper.
-To write out the quantum error scale $\epsilon_V^2$ explicitly, we define a matrix of structure constants
+The remainder of this section is devoted to defining the quantum error scale $\epsilon_V^2$ (derived in Appendix \ref{sec:error}), and can be skipped without consequence for the discussions in the rest of this paper.
+To write out $\epsilon_V^2$ explicitly, we define a matrix of structure constants
 \begin{align}
@@
 \begin{align}
-  \ket{\chi_{VL}}
-  &\equiv \sum_{M=-L}^L \chi^V_{LM} \ket{M}, \\
-  \chi^V_{Lm} &\equiv \sum_\ell \obk{\N_{V\ell}|\D_M|\I g_{L\ell}}, \\
+  \ket{\chi_{V\ell}}
+  &\equiv \sum_{m=-\ell}^\ell \chi^V_{\ell m} \ket{m}, \\
+  \chi^V_{LM} &\equiv \sum_\ell \obk{\N_{V\ell}|\D_M|\I g_{L\ell}}, \\
   \D_M &\equiv \sum_{m,m'} \delta_{M,m'-m} \op{mm'}.
@@
 Here $\D_M$ simply picks off the $M$-th diagonal of the matrix it acts on, such that $\obk{\N_{V\ell}|\D_M|\I g_{L\ell}}$ is an inner product between the $M$-th diagonals of $\N_{V\ell}$ and $\I g_{L\ell}$.
-We then define
-\begin{align}
-  \L_{V\ell} \equiv \N_{V\ell} + \I\N_{V\ell}\I,
-\end{align}
-in terms of which we have that $\epsilon_V^2 \equiv \sum_\ell \epsilon_{V\ell}^2$, with
+The vectors $\ket{\chi_{V\ell}}$ allow us to express reconstruction error in the simple form
+\begin{align}
+  \Bbk{\norm{\tilde\rho_V-\rho}^2}
+  = \f1n \sum_\ell \sp{\bk{\chi_{V\ell}|\rho_\ell}
+    - \bk{\rho_\ell|\N_{V\ell}|\rho_\ell}},
+\end{align}
+where $\ket{\rho_\ell}\equiv\sum_m \rho_{\ell m}\ket{m}$.
+This error is straightforward to maximize with respect to all states $\rho$, which yields the bound in Eq.~\eqref{eq:bound} with $\epsilon_V^2 \equiv \sum_\ell \epsilon_{V\ell}^2$ and
 \begin{align}
@@
   \bk{\chi_{V\ell}|\p{\L_{V\ell}^{-1}
-      - \L_{V\ell}^{-1}\N_{V\ell}\L_{V\ell}^{-1}}|\chi_{V\ell}}.
+      - \L_{V\ell}^{-1}\N_{V\ell}\L_{V\ell}^{-1}}|\chi_{V\ell}},
+\end{align}
+where
+\begin{align}
+  \L_{V\ell} \equiv \N_{V\ell} + \I\N_{V\ell}\I.
 \end{align}

### 2020-11-18 sun_phases/spin_qudit_tomo.tex
enforce trace-1 for the state maximizing reconstruction error

@@
 \end{align}
-where $\bk{X}_\rho\equiv\tr\p{\rho X}$, and $\rho^\dag=\rho$ implies that $\rho_{\ell m}^*=\p{-1}^m\rho_{\ell,-m}$.
+where $\bk{X}_\rho\equiv\tr\p{\rho X}=\obk{\rho|X}$, and $\rho^\dag=\rho$ implies that $\rho_{\ell m}^*=\p{-1}^m\rho_{\ell,-m}$.
 The transition operators can be physically interpreted as the transition $\ket\psi\to T_{\ell m}\ket\psi$ induced on the state $\ket\psi$ of a spin-$s$ system by the absorption of a spin-$\ell$ particle with spin projection $m$ onto a fixed quantization axis.
@@
 \begin{align}
-  \rho_{\ell m} = \bk{T_{\ell m}^\dag}_\rho
-  = \sum_{\v v,k} C_{\ell mk}\p{\v v} \bk{T_{\v vk,0}^\dag}_\rho.
+  \rho_{\ell m}^* = \bk{T_{\ell m}}_\rho
+  = \sum_{\v v,k} C_{\ell mk}\p{\v v} \bk{T_{\v vk,0}}_\rho.
   \label{eq:state_recon}
@@
 \begin{align}
-  T_{\ell m}^\dag
+  T_{\ell m}
   = \sum_{\v v,k,n} C_{\ell mk}\p{\v v}
-  D^k_{n,0}\p{\v v}^* T_{kn}^\dag.
+  D^k_{n,0}\p{\v v} T_{kn}.
 \end{align}
@@
 \begin{align}
-  \sum_{\v v} C_{\ell m}\p{\v v} D^\ell_{n,0}\p{\v v}^* = \delta_{mn}
+  \sum_{\v v} C_{\ell m}\p{\v v} D^\ell_{n,0}\p{\v v} = \delta_{mn}
   \label{eq:tomo_recovery}
@@
 We will refer to Eq.~\eqref{eq:full_recovery} as the {\it full recovery problem}, and Eq.~\eqref{eq:tomo_recovery} as the the {\it reduced recovery problem}.
-Due to the fact that $D^k_{n,0}\p{\v v}^* = \sqrt{\frac{4\pi}{2\ell+1}}\, Y_{kn}\p{\v v}$, any solution to the full recovery problem automatically solves the reduced recovery problem by setting $C_{\ell m}\p{\v v} = \sqrt{\frac{2\ell+1}{4\pi}}\, \tilde C_{\ell m}\p{\v v}$.
+Due to the fact that $D^k_{n,0}\p{\v v} = \sqrt{\frac{4\pi}{2\ell+1}}\, Y_{kn}\p{\v v}^*$, any solution to the full recovery problem automatically solves the reduced recovery problem by setting $C_{\ell m}\p{\v v} = \sqrt{\frac{2\ell+1}{4\pi}}\, \tilde C_{\ell m}\p{\v v}^*$.
 In principle, this mapping allows us to import a host of existing signal recovery algorithms \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality} for the task of spin qudit tomography.
@@
 \end{align}
-where $\norm{X} \equiv \sqrt{\obk{X|X}} = \sqrt{\tr\p{X^\dag X}}$ is the (Hilbert-Schmidt, or Frobenius) norm of $X$, and the {\it classical error scale}
+where $\norm{X}^2 \equiv \obk{X|X} = \tr\p{X^\dag X}$ is the squared (Frobenius, or Hilbert-Schmidt) norm of $X$, and the {\it classical error scale}
 \begin{align}
@@
 For comparison with the bound in Eq.~\eqref{eq:bound_eps}, we note that $\epsilon_V^2<\S_V^2/2$, so the previous bound still holds with the replacement $\epsilon^2\to1/2n$.
-We also note that the (squared) reconstruction error in Eq.~\eqref{eq:error} and its bound in Eq.~\eqref{eq:bound} obey the ``standard quantum limit'' of $\sim1/\sqrt{n}$ scaling in the number of measurements.
-In principle, this scaling can be improved to $\sim1/n$ by preparing and measuring entangled copies of many qudits \cite{giovannetti2006quantum}.
+We also note that the (squared) reconstruction error in Eq.~\eqref{eq:error} and its bound in Eq.~\eqref{eq:bound} obey the ``standard quantum limit'' of $\sim1/n$ scaling in the number of measurements.
+In principle, this scaling can be improved to $\sim1/n^2$ by preparing and measuring entangled copies of many qudits \cite{giovannetti2006quantum}.
 
@@
 The remainder of this section is devoted to defining the quantum error scale $\epsilon_V^2$ (derived in Appendix \ref{sec:error}), and can be skipped without consequence for the discussions in the rest of this paper.
-To write out $\epsilon_V^2$ explicitly, we define a matrix of structure constants
-\begin{align}
-  g_{L\ell} \equiv \sum_{m,n=-\ell}^\ell
-  \obk{T_{L,m+n}|T_{\ell m}T_{\ell n}} \op{m}{n},
-\end{align}
-and the inversion operator $\I\equiv\sum_m\p{-1}^m\op{-m}{m}$, which are used to define the $\p{2\ell+1}$-component vectors
-\begin{align}
-  \ket{\chi_{V\ell}}
-  &\equiv \sum_{m=-\ell}^\ell \chi^V_{\ell m} \ket{m}, \\
-  \chi^V_{LM} &\equiv \sum_\ell \obk{\N_{V\ell}|\D_M|\I g_{L\ell}}, \\
-  \D_M &\equiv \sum_{m,m'} \delta_{M,m'-m} \op{mm'}.
-\end{align}
-Here $\D_M$ simply picks off the $M$-th diagonal of the matrix it acts on, such that $\obk{\N_{V\ell}|\D_M|\I g_{L\ell}}$ is an inner product between the $M$-th diagonals of $\N_{V\ell}$ and $\I g_{L\ell}$.
-The vectors $\ket{\chi_{V\ell}}$ allow us to express reconstruction error in the simple form
+To write out this error in full, we first rewrite the reconstruction error in Eq.~\eqref{eq:error} as
 \begin{align}
@@
     - \bk{\rho_\ell|\N_{V\ell}|\rho_\ell}},
-\end{align}
-where $\ket{\rho_\ell}\equiv\sum_m \rho_{\ell m}\ket{m}$.
-This error is straightforward to maximize with respect to all states $\rho$, which yields the bound in Eq.~\eqref{eq:bound} with $\epsilon_V^2 \equiv \sum_\ell \epsilon_{V\ell}^2$ and
-\begin{align}
-  \epsilon_{V\ell}^2 \equiv
-  \bk{\chi_{V\ell}|\p{\L_{V\ell}^{-1}
-      - \L_{V\ell}^{-1}\N_{V\ell}\L_{V\ell}^{-1}}|\chi_{V\ell}},
-\end{align}
-where
-\begin{align}
-  \L_{V\ell} \equiv \N_{V\ell} + \I\N_{V\ell}\I.
-\end{align}
+  \label{eq:error_simp}
+\end{align}
+where $\ket{\rho_\ell}\equiv\sum_m \rho_{\ell m}\ket{m}$, and similarly $\ket{\chi_{V\ell}}\equiv\sum_m \chi^V_{\ell m} \ket{m}$ is a vector of the scalars $\chi^V_{\ell m}$ defined by
+\begin{align}
+  \chi^V_{LM} &\equiv
+  \sum_\ell \obk{\N_{V\ell}|\D_M|\I g_{L\ell}}, \\
+  g_{L\ell} &\equiv \sum_{m,m'}
+  \obk{T_{L,m+m'}|T_{\ell m}T_{\ell m'}} \op{m}{m'}, \\
+  \I &\equiv \sum_m \p{-1}^m \op{-m}{m}, \\
+  \D_M &\equiv \sum_{m,m'} \delta_{M,m'-m} \op{mm'}.
+\end{align}
+Here $g_{L\ell}$ is essentially a matrix of transition operator product coefficients (see Appendix \ref{sec:trans_prod}), and $\D_M$ simply picks off the $M$-th diagonal of the matrix it acts on.
+Maximizing the reconstruction error in Eq.~\eqref{eq:error_simp} over all $\rho$ with $\tr\rho=1$ leads to the bound in Eq.~\eqref{eq:bound} with
+\begin{align}
+  \epsilon_V^2 \equiv \sum_\ell\sp{\bk{\chi_{V\ell}|\rho_{V\ell}^\star}
+    - \bk{\rho_{V\ell}^\star|\N_{V\ell}|\rho_{V\ell}^\star}}
+\end{align}
+where the maximizers
+\begin{align}
+  \ket{\rho_{V\ell}^\star} \stackrel{\ell>0}{\equiv}
+  \L_{V\ell}^{-1} \ket{\chi_{V\ell}},
+  &&
+  \ket{\rho_{V,0}^\star} \equiv \f1{\sqrt{d}} \ket{0},
+\end{align}
+are defined in terms of the inverse $\L_{V\ell}^{-1}$ of
+\begin{align}
+  \L_{V\ell} \equiv \N_{V\ell} + \I \N_{V\ell} \I.
+\end{align}
+Note that the bound in Eq.~\eqref{eq:bound} is obtained by maximizing the reconstruction error in Eq.~\eqref{eq:error_simp} over all possible $\rho$ (satisfying $\tr\rho=1$) regardless of their spectrum.
+As a consequence, the maximizing ``state'' $\rho_V^\star$ may generally have negative eigenvalues.
+In principle, there is therefore room for improvement on the bound in Eq.~\eqref{eq:bound} by maximizing reconstruction error over the (restricted) domain of states with a non-negative spectrum.
 
@@
   = \f1n \sum_{i,\ell,m,m'}
-  \abs{\tilde{\v q}_{\ell i}}^2
-  D^\ell_{m,0}\p{\v v_i}^* D^\ell_{m',0}\p{\v v_i}
+  D^\ell_{m',0}\p{\v v_i} \abs{\tilde{\v q}_{\ell i}}^2
+  D^\ell_{m,0}\p{\v v_i}^*
   \cov_\rho\p{T_{\ell m}^\dag, T_{\ell m'}}.
@@
   \Bbk{\norm{\tilde\rho_V-\rho}^2}
-  = \f1n \sum_\ell \obk{\N_{V\ell}|\C_\ell\sp{\rho}}.
+  = \f1n \sum_\ell \obk{\N_{V\ell}|\C_\ell\sp{\rho}},
   \label{eq:error_SM}
 \end{align}
+where $\obk{X|Y}=\tr\p{X^\dag Y}$ is a trace inner product.
 
@@
   - \sum_{\ell,m,m'} \bk{m'|\I\N_{V\ell}|m}
-  \rho_{\ell m'} \rho_{\ell m},
+  \rho_{\ell m} \rho_{\ell m'},
   \label{eq:error_comps}
@@
 Here $\D_M$ simply picks off the $M$-th diagonal of the matrix it acts on.
-In total, reconstruction error can be written in the form
-\begin{align}
-  n \Bbk{\norm{\tilde\rho_V-\rho}^2}
-  = \obk{\chi_V|\rho}
-  - \sum_\ell \obk{\rho|U_\ell\N_{V\ell}U_\ell^\dag|\rho},
-  &&
-  \oket{\chi_V} \equiv \sum_{L,M} \chi^V_{LM} \oket{T_{LM}},
-  &&
-  U_\ell \equiv \sum_m \oket{T_{\ell m}} \bra{m}.
-\end{align}
-Performing the change of basis $\oket{T_{\ell m}}\to\ket{\ell m}$ by defining
-\begin{align}
-  \ket{\rho} \equiv U^\dag \oket{\rho}
-  = \sum_{\ell,m} \rho_{\ell m}\ket{\ell m},
-  &&
-  \ket{\chi_V} \equiv U^\dag \oket{\chi_V}
-  = \sum_{\ell,m} \chi^V_{\ell m} \ket{\ell m},
-  &&
-  U \equiv \sum_{\ell,m} \oket{T_{\ell m}} \bra{\ell m},
-\end{align}
-as well as the degree-$\ell$ projections
+Defining the $\p{2\ell+1}$-component vectors
 \begin{align}
@@
   &&
-  \ket{\chi_{V\ell}} \equiv \sum_m \chi^V_{\ell m} \ket{m},
-\end{align}
-we can more simply write
+  \ket{\chi_{V\ell}} \equiv \sum_{\ell,m} \chi^V_{\ell m} \ket{m},
+\end{align}
+the expansion in Eq.~\eqref{eq:error_comps} is most efficiently evaluated in the vectorized form
 \begin{align}
@@
 \end{align}
-Given the expansion in Eq.~\eqref{eq:error_comps}, this error is straightforward to maximize by considering its variation with respect to $\rho_{\ell m}$:
+This error is straightforward to maximize by considering its variation with respect to $\rho_{\ell m}$:
 \begin{align}
@@
   = \p{\chi^V_{\ell m}}^*
-  - \sum_{m'} \sp{\bk{m'|\I\N_{V\ell}|m} + \bk{m'|\I\N_{V\ell}|m}}
+  - \sum_{m'} \sp{\bk{m'|\I\N_{V\ell}|m} + \bk{m|\I\N_{V\ell}|m'}}
   \rho_{\ell m'}.
@@
 \begin{align}
-  \ket{\chi_{V\ell}} - \L_{V\ell} \ket{\rho_{V\ell}^\star} = 0,
+  \ket{\chi_{V\ell}} - \L_{V\ell} \ket{\rho_{V\ell}^\star}
+  \stackrel{\ell>0}{=} 0,
   &&
@@
 \begin{align}
-  \ket{\rho_{V\ell}^\star} = \L_{V\ell}^{-1} \ket{\chi_{V\ell}},
-\end{align}
-and in turn
+  \ket{\rho_{V\ell}^\star} \stackrel{\ell>0}{=}
+  \L_{V\ell}^{-1} \ket{\chi_{V\ell}},
+  &&
+  \ket{\rho_{V,0}^\star} = \f1{\sqrt{d}} \ket{0},
+\end{align}
+where the $\ell=0$ component of $\rho_V^\star$ enforces $\tr\rho_V^\star=1$.
+We thus find that
 \begin{align}
@@
   \le \Bbk{\norm{\tilde\rho_V-\rho_V^\star}^2}
-  = \f{\epsilon_V^2}{n},
+  = \f1n \sum_\ell \epsilon_{V\ell}^2,
+  &&
+  \epsilon_{V\ell}^2 \equiv \bk{\chi_{V\ell}|\rho_{V\ell}^\star}
+  - \bk{\rho_{V\ell}^\star|\N_{V\ell}|\rho_{V\ell}^\star}.
   \label{eq:bound_SM}
 \end{align}
-where
-\begin{align}
-  \epsilon_V^2 \equiv \sum_\ell \epsilon_{V\ell}^2,
-  &&
-  \epsilon_{V\ell}^2 \equiv \bk{\chi_{V\ell}|\p{\L_{V\ell}^{-1}
-      - \L_{V\ell}^{-1}\N_{V\ell}\L_{V\ell}^{-1}}|\chi_{V\ell}}.
-\end{align}
-Note that the bound in Eq.~\eqref{eq:bound_SM} is obtained by maximizing reconstruction error over all possible $\rho$ regardless of their spectrum, which implies that the maximizer $\rho_V^\star$ may generally be an unphysical state with negative eigenvalues.
+Note that the bound in Eq.~\eqref{eq:bound_SM} is obtained by maximizing reconstruction error over all possible $\rho$ (satisfying $\tr\rho=1$) regardless of their spectrum, which implies that the maximizer $\rho_V^\star$ may generally be an unphysical ``state'' with negative eigenvalues.
 In principle, there is therefore room for improvement on the bound in Eq.~\eqref{eq:bound_SM} by maximizing reconstruction error over the (restricted) domain of states with a non-negative spectrum.

### 2020-11-18 sun_phases/spin_qudit_tomo.tex
simplified reconstruction error maximum
minor changes to reconsturction error maximization work
minor changes to text corrected a typo

@@
 \begin{align}
-  \epsilon_V^2 \equiv \sum_\ell\sp{\bk{\chi_{V\ell}|\rho_{V\ell}^\star}
-    - \bk{\rho_{V\ell}^\star|\N_{V\ell}|\rho_{V\ell}^\star}}
-\end{align}
-where the maximizers
-\begin{align}
-  \ket{\rho_{V\ell}^\star} \stackrel{\ell>0}{\equiv}
-  \L_{V\ell}^{-1} \ket{\chi_{V\ell}},
-  &&
-  \ket{\rho_{V,0}^\star} \equiv \f1{\sqrt{d}} \ket{0},
-\end{align}
-are defined in terms of the inverse $\L_{V\ell}^{-1}$ of
-\begin{align}
-  \L_{V\ell} \equiv \N_{V\ell} + \I \N_{V\ell} \I.
-\end{align}
-Note that the bound in Eq.~\eqref{eq:bound} is obtained by maximizing the reconstruction error in Eq.~\eqref{eq:error_simp} over all possible $\rho$ (satisfying $\tr\rho=1$) regardless of their spectrum.
-As a consequence, the maximizing ``state'' $\rho_V^\star$ may generally have negative eigenvalues.
-In principle, there is therefore room for improvement on the bound in Eq.~\eqref{eq:bound} by maximizing reconstruction error over the (restricted) domain of states with a non-negative spectrum.
+  \epsilon_V^2 \equiv \sum_{\ell>0} \sp{\f14 \bk{\chi_{V\ell}|\N_{V\ell}^{-1}|\chi_{V\ell}}
+    + \f1d \tr\p{\N_{V\ell}}}.
+\end{align}
+Note that the maximum of Eq.~\eqref{eq:error_simp} may generally occur at an unphysical ``state'' $\rho$ with negative eigenvalues.
+In principle, there is therefore room for improvement on the bound in Eq.~\eqref{eq:bound} by maximizing reconstruction error over the restricted domain of states with a non-negative spectrum.
 
@@
 \begin{align}
-  \ket{\chi_{V\ell}} - \L_{V\ell} \ket{\rho_{V\ell}^\star}
+  \ket{\chi_{V\ell}}
+  - \sp{\N_{V\ell} + \I N_{V\ell}^* \I} \ket{\rho_{V\ell}^\star}
+  = \ket{\chi_{V\ell}}
+  - 2 \N_{V\ell} \ket{\rho_{V\ell}^\star}
   \stackrel{\ell>0}{=} 0,
-  &&
-  \L_{V\ell} \equiv \N_{V\ell} + \I N_{V\ell} \I,
 \end{align}
@@
   \ket{\rho_{V\ell}^\star} \stackrel{\ell>0}{=}
-  \L_{V\ell}^{-1} \ket{\chi_{V\ell}},
+  \f12 \N_{V\ell}^{-1} \ket{\chi_{V\ell}},
   &&
@@
   \le \Bbk{\norm{\tilde\rho_V-\rho_V^\star}^2}
-  = \f1n \sum_\ell \epsilon_{V\ell}^2,
-  &&
-  \epsilon_{V\ell}^2 \equiv \bk{\chi_{V\ell}|\rho_{V\ell}^\star}
-  - \bk{\rho_{V\ell}^\star|\N_{V\ell}|\rho_{V\ell}^\star}.
+  = \f{\epsilon_V^2}{n},
   \label{eq:bound_SM}
 \end{align}
+where
+\begin{align}
+  \epsilon_V^2
+  \equiv \sum_{\ell>0} \f14 \bk{\chi_{V\ell}|\N_{V\ell}^{-1}|\chi_{V\ell}}
+  + \sp{\f{\bk{\chi_{V,0}|0}}{\sqrt{d}} - \f{\bk{0|\N_{V,0}|0}}{d}}
+  = \sum_{\ell>0} \sp{\f14 \bk{\chi_{V\ell}|\N_{V\ell}^{-1}|\chi_{V\ell}}
+    + \f1d \tr\p{\N_{V\ell}}}.
+\end{align}
 Note that the bound in Eq.~\eqref{eq:bound_SM} is obtained by maximizing reconstruction error over all possible $\rho$ (satisfying $\tr\rho=1$) regardless of their spectrum, which implies that the maximizer $\rho_V^\star$ may generally be an unphysical ``state'' with negative eigenvalues.
-In principle, there is therefore room for improvement on the bound in Eq.~\eqref{eq:bound_SM} by maximizing reconstruction error over the (restricted) domain of states with a non-negative spectrum.
+In principle, there is therefore room for improvement on the bound in Eq.~\eqref{eq:bound_SM} by maximizing reconstruction error over the restricted domain of states with a non-negative spectrum.
 

### 2020-11-19 sun_phases/spin_qudit_tomo.tex
gave up on a tight reconstruction error bound due to difficulties with maximizing over the space of physical states with a non-negative spectrum

@@
 \newcommand{\D}{\mathcal{D}}
+\newcommand{\E}{\mathcal{E}}
 \newcommand{\I}{\mathcal{I}}
@@
 
-[\red{todo: comment on connection between $T_{\ell m}$ and $Y_{\ell m}$}]
+[\red{todo: investigate and comment on the connection between $T_{\ell m}$ and $Y_{\ell m}$?}]
 
@@
 
-[\red{todo: add visualization of the signal recovery problem}]
+[\red{todo: add figure for signal recovery problem?}]
 
@@
 Unfortunately, we have no general proof of this conjecture, which requires a deeper analysis of fixed-degree signal recovery on the sphere.
-Nonetheless, in the following section we numerically verify that $2d-1$ measurement axes suffice for all $d\le200$ [\red{todo: possibly refer to Section \ref{sec:axes} instead}].
+Nonetheless, in the following section we numerically verify that $2d-1$ measurement axes suffice for all $d\le200$ [\red{todo: refer to Section \ref{sec:axes} instead}].
 
@@
 \end{align}
-where for shorthand we use a combined index $j=\p{\v v_j,\ell_j}$ to specify both a measurement axis $\v v_j$ and a degree $\ell_j$, which identify the transition operator $T_j\equiv T_{\v v_j\ell_j,0}$.
+where for shorthand we use a combined index $j=\p{\v v,\ell}$ to specify both a measurement axis $\v v$ and a degree $\ell$, which identify the transition operator $T_j\equiv T_{\v v\ell,0}$.
 These operators allow us to expand any state $\rho$ of a $d$-level spin qudit in the form
@@
 \end{align}
-where $\set{\epsilon_j}$ are independent random variables and we use the double brackets $\bbk{\cdot}$ to denote statistical averaging.
+where $\set{\epsilon_j}$ are independent random variables and we use the double brackets $\bbk{\cdot}$ to denote statistical averaging over experimental trials that estimate $\bk{T_j}_\rho$.
 In this case, the mean squared error with which $\tilde\rho^V_k$ approximates $\rho^V_k$ is
@@
 \begin{align}
-  \Bbk{\norm{\tilde\rho_V-\rho}^2}
+  \E_V\p{\rho}^2
+  \equiv \Bbk{\norm{\tilde\rho_V-\rho}^2}
   \le \epsilon^2 \S_V^2,
@@
 \end{align}
-is the squared norm of the right inverse $M_V^{-1}$ of $M_V$, such that $M_V M_V^{-1}=1$.
-Note that the classical error scale $\S_V^2$ diverges if $M_V$ is singular, in which case measuring spin projections along all axes in $V$ does not provide sufficient information to reconstruct arbitrary quantum states.
+is the squared norm of the right inverse $M_V^{-1}$ of $M_V$, satisfying $M_V M_V^{-1}=1$.
+We refer to this error scale $\S_V^2$ as ``classical'' because the bound in Eq.~\eqref{eq:bound_eps} applies in the presence of classical sources of measurement error.
+Note that the classical error scale $\S_V^2$ diverges if the measurement matrix $M_V$ is singular, in which case measuring spin projections along all axes in $V$ does not provide sufficient information to reconstruct arbitrary quantum states.
 
@@
 \end{align}
-where the $\p{2\ell+1}\times\abs{V}$ blocks are
+where the $\p{2\ell+1}\times\abs{V}$-sized blocks are
 \begin{align}
@@
   \includegraphics{error_scale/error_times.pdf}
-  \caption{Serial runtime to compute the error scales $\S_V^2$ and $\epsilon_V^2$ with $\abs{V}=2d-1$ randomly chosen measurement axes.
+  \caption{Serial runtime to compute the classical error scale $\S_V^2$ with $\abs{V}=2d-1$ randomly chosen measurement axes.
     Each point is an average over 100 calculations or 5 minutes of runtime, whichever comes first.
@@
 However, this assumption breaks down when measurement error is limited by fundamental quantum (shot) noise.
-We relax the assumption of Eq.~\eqref{eq:error_assumption} in Appendix \ref{sec:error}, where we instead assume that $\tilde\rho_V$ is built from $n$ independent measurements of spin projection along every axis $\v v\in V$, with shot noise the dominant source of error.
-In this case, the mean squared distance between the approximation $\tilde\rho_V$ and the true state $\rho$ can be compactly expressed using trace inner products of the {\it covariance matrices}
-\begin{align}
-  \C_\ell\sp{\rho} \equiv \sum_{m,n=-\ell}^\ell
-  \cov_\rho\p{T_{\ell m}^\dag, T_{\ell n}} \op{m}{n}
-\end{align}
-where $\cov_\rho\p{X,Y} \equiv \bk{XY}_\rho - \bk{X}_\rho \bk{Y}_\rho$, and the {\it noise matrices}
+We relax the assumption of Eq.~\eqref{eq:error_assumption} in Appendix \ref{sec:improved_bound}, where we instead assume that $\tilde\rho_V$ is built from $n$ independent measurements of spin projection along every axis $\v v\in V$, with shot noise the dominant source of error.
+In this case, the constraints that all $\bk{\Pi_{\v v\mu}}_\rho\ge0$ and $\tr\p\rho=1$ allow us to bound the mean squared distance between $\tilde\rho_V$ and $\rho$ as
+\begin{align}
+  \E_V\p{\rho}^2 < \f{\epsilon_V^2}{n},
+  &&
+  \epsilon_V^2 \equiv \sum_\ell \Gamma_\ell^2 \S_{V\ell}^2,
+  \label{eq:bound}
+\end{align}
+where the {\it quantum error scale} $\epsilon_V^2$ is defined in terms of (half of) the spectral range of $T_{\ell,0}$:
+\begin{align}
+  \Gamma_\ell \equiv \f{\max_\mu t_{\ell\mu} - \min_\mu t_{\ell\mu}}{2},
+  &&
+  t_{\ell\mu} \equiv \bk{\mu|T_{\ell,0}|\mu}.
+\end{align}
+If $d$ is even, then $\Gamma_\ell=\max_\mu t_{\ell\mu}$.
+For comparison with previous error bound in Eq.~\eqref{eq:bound_eps}, we note that $\epsilon_V^2<\S_V^2/2$, so the previous bound still holds with the replacement $\epsilon^2\to1/2n$.
+The factors $\Gamma_\ell^2$ are quick to compute and can be recycled for every new choice of axes $V$, so the complexity of computing $\epsilon_V^2$ is identical to that of $\S_V^2$.
+
+Though straightforward to compute, the bound in Eq.~\eqref{eq:bound} is not tight, as it is acquired by bounding the statistical variance $\epsilon_{\v v\ell}^2$ in the empirical estimate $\tilde T_{\v v\ell,0}$ of $\bk{T_{\v v\ell,0}}_\rho$ as $\bbk{\epsilon_{\v v\ell}^2}\le\Gamma_\ell^2$.
+Though the individual bounds on $\bbk{\epsilon_{\v v\ell}^2}$ for each axis $\v v$ and degree $\ell$ are tight, these bounds cannot all be achieved simultaneously.
+There is therefore still room for improvement on the bound in Eq.~\eqref{eq:bound} by maximizing $\E_V$ over the set of all qudit states $\rho$.
+We also note that the (squared) reconstruction error bound in Eq.~\eqref{eq:bound} obeys the ``standard quantum limit'' of $\sim1/n$ scaling in the number of measurements.
+In principle, this scaling can be improved to $\sim1/n^2$ by preparing and measuring entangled copies of many qudits \cite{giovannetti2006quantum}.
+
+The quantum error scale $\epsilon_V^2$ provides a pessimistic upper bound on statistical error, which can be calculated without prior knowledge of the true qudit state $\rho$.
+The actual error in the reconstruction $\tilde\rho_V$ of a particular state $\rho$ may be considerably smaller.
+The exact mean squared distance between $\tilde\rho_V$ and $\rho$, derived in Appendix \ref{sec:exact}, can be written in the form
+\begin{align}
+  \E_V\p{\rho}^2
+  = \f1n \sum_\ell \sp{\bk{\chi_{V\ell}|\rho_\ell}
+    - \bk{\rho_\ell|\N_{V\ell}|\rho_\ell}},
+  \label{eq:error}
+\end{align}
+where $\ket{\rho_\ell} \equiv \sum_m \rho_{\ell m}\ket{m}$, and $\ket{\chi_{V\ell}}$ and $\N_{V\ell}$ are defined below.
+While the true error in $\tilde\rho_V$ cannot be known exactly without knowing $\rho$, this error can be estimated by computing the right-hand side of Eq.~\eqref{eq:error} with the replacement $\rho\to\tilde\rho_V$.
+
+The matrix $\N_{V\ell}$ that appears in Eq.~\eqref{eq:error} is defined in terms of the measurement matrix $M_{V\ell}$ and its right inverse $M_{V\ell}^{-1}$ by
 \begin{align}
@@
 \end{align}
-where $\diag\sp{X}$ sets all off-diagonal parts of $X$ to zero:
-\begin{align}
-  \Bbk{\norm{\tilde\rho_V-\rho}^2}
-  = \f1n \sum_\ell \obk{\N_{V\ell}|\C_\ell\sp{\rho}}.
-  \label{eq:error}
-\end{align}
-Maximizing this error over all $\rho$ (for a fixed set of axes $V$) yields a reconstruction error bound of the form
-\begin{align}
-  \Bbk{\norm{\tilde\rho_V-\rho}^2} \le \f{\epsilon_V^2}{n},
-  \label{eq:bound}
-\end{align}
-where the complexity of computing the {\it quantum error scale} $\epsilon_V^2$ (defined below) is the same as that of computing the classical error scale $\S_V^2$ (see Figure \ref{fig:times}).
-For comparison with the bound in Eq.~\eqref{eq:bound_eps}, we note that $\epsilon_V^2<\S_V^2/2$, so the previous bound still holds with the replacement $\epsilon^2\to1/2n$.
-We also note that the (squared) reconstruction error in Eq.~\eqref{eq:error} and its bound in Eq.~\eqref{eq:bound} obey the ``standard quantum limit'' of $\sim1/n$ scaling in the number of measurements.
-In principle, this scaling can be improved to $\sim1/n^2$ by preparing and measuring entangled copies of many qudits \cite{giovannetti2006quantum}.
-
-The quantum error scale $\epsilon_V^2$ only provides a pessimistic upper bound on statistical error, which can be calculated without prior knowledge of the true qudit state $\rho$.
-The actual error in the reconstruction $\tilde\rho_V$ of a particular state $\rho$, provided by Eq.~\eqref{eq:error}, may be considerably smaller.
-While the true error in $\tilde\rho_V$ cannot be known exactly without knowing $\rho$, this error can be estimated by computing the right-hand side of Eq.~\eqref{eq:error} with the replacement $\rho\to\tilde\rho_V$.
-
-The remainder of this section is devoted to defining the quantum error scale $\epsilon_V^2$ (derived in Appendix \ref{sec:error}), and can be skipped without consequence for the discussions in the rest of this paper.
-To write out this error in full, we first rewrite the reconstruction error in Eq.~\eqref{eq:error} as
-\begin{align}
-  \Bbk{\norm{\tilde\rho_V-\rho}^2}
-  = \f1n \sum_\ell \sp{\bk{\chi_{V\ell}|\rho_\ell}
-    - \bk{\rho_\ell|\N_{V\ell}|\rho_\ell}},
-  \label{eq:error_simp}
-\end{align}
-where $\ket{\rho_\ell}\equiv\sum_m \rho_{\ell m}\ket{m}$, and similarly $\ket{\chi_{V\ell}}\equiv\sum_m \chi^V_{\ell m} \ket{m}$ is a vector of the scalars $\chi^V_{\ell m}$ defined by
+where $\diag\sp{X}$ sets all off-diagonal parts of $X$ to zero.
+The vector $\ket{\chi_{V\ell}}\equiv\sum_m \chi^V_{\ell m} \ket{m}$ in Eq.~\eqref{eq:error} is defined by
 \begin{align}
@@
 \end{align}
-Here $g_{L\ell}$ is essentially a matrix of transition operator product coefficients (see Appendix \ref{sec:trans_prod}), and $\D_M$ simply picks off the $M$-th diagonal of the matrix it acts on.
-Maximizing the reconstruction error in Eq.~\eqref{eq:error_simp} over all $\rho$ with $\tr\rho=1$ leads to the bound in Eq.~\eqref{eq:bound} with
-\begin{align}
-  \epsilon_V^2 \equiv \sum_{\ell>0} \sp{\f14 \bk{\chi_{V\ell}|\N_{V\ell}^{-1}|\chi_{V\ell}}
-    + \f1d \tr\p{\N_{V\ell}}}.
-\end{align}
-Note that the maximum of Eq.~\eqref{eq:error_simp} may generally occur at an unphysical ``state'' $\rho$ with negative eigenvalues.
-In principle, there is therefore room for improvement on the bound in Eq.~\eqref{eq:bound} by maximizing reconstruction error over the restricted domain of states with a non-negative spectrum.
+Here $g_{L\ell}$ is a matrix of structure constants for the transition operator algebra (see Appendix \ref{sec:trans_prod}), and $\D_M$ simply picks off the $M$-th diagonal of the matrix it acts on, such that $\obk{\N_{V\ell}|\D_M|\I g_{L\ell}}$ is an inner product of the $M$-th diagonal of $\I g_{L\ell}$ with the $\p{-M}$-th diagonal of $\N_{V\ell}$.
+
+[\red{todo: comment on code availability}]
 
@@
 
-The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and optimize (using any standard minimization algorithm) the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the quantum error scale $\epsilon_V^2$.
+The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and optimize (using any standard minimization algorithm) the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the quantum error scale $\epsilon_V^2$ defined in Eq.~\eqref{eq:bound}.
 If $\abs{V}$ is too large for such optimization to be practical, one can instead generate many random sets of measurement axes, and then choose the set with the smallest quantum error scale $\epsilon_V^2$.
 After choosing a set of measurement axes $V$, (experimentally) make $n$ measurements of spin projection along every axis $\v v\in V$, and set $\tilde\Pi_{\v v\mu}\approx\bk{\Pi_{\v v\mu}}_\rho$ to the fraction of times in which the measurement outcome was $\mu$.
-The estimates $\tilde\Pi_{\v v\mu}$ of $\bk{\Pi_{\v v\mu}}_\rho$ determine estimates of the expectation values $\bk{T_{\v v\ell,0}}_\rho$, which can be used to construct an estimate $\tilde\rho_V$ of $\rho$.
+The estimates $\tilde\Pi_{\v v\mu}$ of $\bk{\Pi_{\v v\mu}}_\rho$ determine estimates of the expectation values $\bk{T_{\v v\ell,0}}_\rho$, which can be used to construct an estimate $\tilde\rho_V$ of $\rho$, defined in Eq.~\eqref{eq:reconstructed_state}.
 If $\tilde\rho_V$ has negative eigenvalues, its distance from $\rho$ can be reduced with maximum-likelihood corrections \cite{smolin2012efficient}.
+Finally, the error in $\tilde\rho_V$ can be estimated by computing the right-hand side of Eq.~\eqref{eq:error} with the replacement $\rho_V\to\tilde\rho_V$.
 
@@
 As a last point, we mention that the quantum error scale $\epsilon_V^2$ can also be used to evaluate the ``quality'' of measurement axes $V$ chosen with existing methods for sparse signal recovery on the sphere \cite{rauhut2011sparse, alem2012sparse}.
-In practice, we find that uniform random sampling outperforms these methods, in the sense that it generates low-error-scale sets of measurement axes with higher probability [\red{todo: add figure}].
+In practice, we find that uniform random sampling outperforms these methods, in the sense that it generates low-error-scale sets of measurement axes with higher probability [\red{todo: add figure?}].
 Two observations may explain the poor performance of these sparse signal recovery methods.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Revisiting the reconstruction error bound}
-\label{sec:error}
+\section{An improved reconstruction error bound}
+\label{sec:improved_bound}
 
@@
 If $\p{i,j}\ne\p{i',j'}$, then $\Pi_{i\mu}^j$ and $\Pi_{i'\mu'}^{j'}$ address different tensor factors of the product state $\rho^{\otimes N}$, so the expectation value of their product factorizes due to the identity $\tr\sp{\p{A\otimes B}\p{A'\otimes B'}} = \tr\p{AA'}\times\tr\p{BB'}$.
-This factorization can also be seen as a consequence of the fact that if $\p{i,j}\ne\p{i',j'}$, then $\Pi_{i\mu}^j$ and $\Pi_{i'\mu'}^{j'}$ are ``spatially separated'', which means that their expectation values with respect to the product state $\rho^{\otimes N}$ cannot have quantum correlations.
+This factorization can also be seen as a consequence of the fact that if $\p{i,j}\ne\p{i',j'}$, then $\Pi_{i\mu}^j$ and $\Pi_{i'\mu'}^{j'}$ are ``spatially separated'' on $\rho^{\otimes N}$, which means that their expectation values cannot have quantum correlations.
 The terms in Eq.~\eqref{eq:proj_cov_start} with $\p{i,j}\ne\p{i',j'}$ therefore vanish, so
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Covariance of errors in the transition operator basis}
+\subsection{Variance of errors in the transition operator basis}
 
@@
 where we used the fact that the covariance $\cov_\rho\p{X,Y}$ is linear in both $X$ and $Y$.
+Due to the appearance of $\delta_{ii'}$ above and the orthogonality of rotated transition operators $T_{i\ell}$ and $T_{i'\ell'}$ with degrees $\ell\ne\ell'$, only the variance $\bbk{\epsilon_{i\ell}^2}$ ultimately contributes to reconstruction error.
+We therefore seek to find an upper bound on $\bbk{\epsilon_{i\ell}^2}$.
+
+To this end, we define the probability $p^i_\mu\equiv\bk{\Pi_{i\mu}}_\rho$, collect these probabilities into the classical probability distribution $p^i = \p{p^i_{-s}, p^i_{-s+1}, \cdots, p^i_s}$, and define the vector $t_\ell \equiv \p{t_{\ell,-s},t_{\ell,-s+1},\cdots,t_{\ell s}}$ of all $t_{\ell\mu}\equiv\bk{\mu|T_{\ell,0}|\mu}$.
+We then observe that
+\begin{align}
+  \bbk{\epsilon_{i\ell}^2} = \f1n \times \sigma_{p^i}^2\p{t_\ell},
+  &&
+  \sigma_p^2\p{X}
+  \equiv \sum_\mu p_\mu X_\mu^2 - \p{\sum_\mu p_\mu X_\mu}^2,
+\end{align}
+where $\sigma_p^2\p{X}$ is the weighted variance of $X$.
+This variance is maximal when $p$ has equal weight on the largest and smallest values of $X$, which implies that
+\begin{align}
+  \sigma_p^2\p{t_\ell} \le \Gamma_\ell^2,
+  &&
+  \Gamma_\ell \equiv \f{\max_\mu t_{\ell\mu} - \min_\mu t_{\ell\mu}}{2},
+  &&
+  \t{so}
+  &&
+  \bbk{\epsilon_{i\ell}^2} \le \f1n \times \Gamma_\ell^2.
+\end{align}
+Note that this bound on $\bbk{\epsilon_{i\ell}^2}$ is tight, as equality is achieved by the state
+\begin{align}
+  \rho_i^\star = \f12\p{\Pi_{i\mu_{\t{max}}} + \Pi_{i\mu_{\t{min}}}}.
+\end{align}
+To find a nicer analytical bound on $\bbk{\epsilon_{i\ell}^2}$, we can use normalization of the transition operators, $\obk{T_{i\ell}|T_{i\ell}}=\sum_\mu t_{\ell\mu}^2 = 1$, and the fact that all probabilities $p_\mu\le1$ to bound
+\begin{align}
+  \sigma_p^2\p{t_\ell} \le \sum_\mu p_\mu t_{\ell\mu}^2
+  \le \sum_\mu t_{\ell\mu}^2 = 1,
+  &&
+  \t{so}
+  &&
+  \bbk{\epsilon_{i\ell}^2} < \f1n.
+\end{align}
+We can get a tighter analytical bound on $\bbk{\epsilon_{i\ell}^2}$ by considering the fact that $t_{\ell\mu}^2 = t_{\ell,-\mu}^2$ due to the symmetries of the Clebsch-Gordan coefficients.
+Denoting the index that maximizes $t_{\ell\mu}^2$ by $\mu_{\t{max}}$, it follows that if $\mu_{\t{max}}\ne0$ then
+\begin{align}
+  \sigma_p^2\p{t_\ell} \le \sum_\mu p_\mu t_{\ell\mu}^2
+  \le t_{\ell\mu_{\t{max}}}^2
+  = \f12 \p{t_{\ell\mu_{\t{max}}}^2 + t_{\ell,-\mu_{\t{max}}}^2}
+  \stackrel{\mu_{\t{max}}\ne0}{\le} \f12 \sum_\mu t_{\ell\mu}^2
+  = \f12.
+\end{align}
+If $\mu_{\t{max}}=0$, then similarly
+\begin{align}
+  t_{\ell\mu_{\t{max}}}^2 + 2 t_{\ell\mu_{\t{min}}}^2
+  = t_{\ell\mu_{\t{max}}}^2 + t_{\ell\mu_{\t{min}}}^2
+  + t_{\ell,-\mu_{\t{min}}}^2
+  \stackrel{\mu_{\t{max}}=0}{\le}
+  \sum_\mu t_{\ell\mu}^2 = 1,
+  &&
+  \t{so}
+  &&
+  \abs{t_{\ell\mu_{\t{min}}}}
+  \stackrel{\mu_{\t{max}}=0}{\le}
+  \sqrt{\f{1 - t_{\ell\mu_{\t{max}}}^2}{2}},
+\end{align}
+where $\mu_{\t{min}}$ is the index that minimizes $t_{\ell\mu}^2$.
+This inequality lets us bound
+\begin{align}
+  \Gamma_\ell
+  = \f12 \p{t_{\ell\mu_{\t{max}}} - t_{\ell\mu_{\t{min}}}}
+  \le \f12 \p{t_{\ell\mu_{\t{max}}} + \abs{t_{\ell\mu_{\t{min}}}}}
+  \stackrel{\mu_{\t{max}}=0}{\le}
+  \f12 t_{\ell\mu_{\t{max}}}
+  + \f12\sqrt{\f{1 - t_{\ell\mu_{\t{max}}}^2}{2}}
+  \equiv \lambda\p{t_{\ell\mu_{\t{max}}}}.
+\end{align}
+If is straightforward to show that $\lambda\p{x}$ is maximally $\lambda^\star\equiv \max_x \lambda\p{x} = \sqrt{3/8}$, so
+\begin{align}
+  \Gamma_\ell^2
+  \stackrel{\mu_{\t{max}}=0}{\le} \p{\lambda^\star}^2
+  = \f38 < \f12.
+\end{align}
+Altogether, we thus find that in all cases
+\begin{align}
+  \sigma_p^2\p{t_\ell} \le \f12,
+  &&
+  \t{so}
+  &&
+  \bbk{\epsilon_{i\ell}} \le \f1{2n}.
+\end{align}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Exact reconstruction error}
-\label{sec:error_exact}
-
-We now revisit the derivation of reconstruction error in Section \ref{sec:bound}.
-To summarize, for a set of measurement axes $V=\set{\v v}$ and degrees $\ell\in\set{0,1,\cdots,d-1}$ we construct the measurement matrix
+\subsection{Revisiting the reconstruction error bound}
+\label{sec:revisiting}
+
+We now revisit the derivation of reconstruction error in Section \ref{sec:bound} to make use of the bounds on variances $\bbk{\epsilon_{i\ell}^2}$.
+To recap, for a set of measurement axes $V=\set{\v v}$ and degrees $\ell\in\set{0,1,\cdots,d-1}$ we construct the measurement matrix
 \begin{align}
@@
 The block-diagonal structure of $M_V$ allows us to index its singular values $M^V_{\ell m}$ and corresponding (normalized) right singular vectors $\v x^V_{\ell m} = (x^V_{\ell m,1}, x^V_{\ell m,2}, \cdots )$ by the indices $\p{\ell,m}$, where the integer $\abs{m}\le\ell$.
-These singular vectors and values allow us to define the orthonormal operators
+These singular vectors and values define the orthonormal operators
 \begin{align}
@@
 where $i\in\set{1,2,\cdots,\abs{V}}$ indexes an axis $\v v_i\in V$.
-The state $\rho$ can be expanded in the basis of $\set{Q^V_{\ell m}}$ as
+The state $\rho$ can be expanded in the basis of these operators as
 \begin{align}
@@
 \end{align}
-Expanding $\tilde T_{i\ell} = \bk{T_{i\ell}}_\rho + \epsilon_{i\ell}$, we can use orthonormality of all $Q^V_{\ell m}$ to expand the mean squared distance between $\tilde\rho_V$ and $\rho$ as
-\begin{align}
-  \Bbk{\norm{\tilde\rho_V-\rho}^2}
+Recalling that $\epsilon_{i\ell} \equiv \tilde T_{i\ell} - \bk{T_{i\ell}}_\rho$, we can use orthonormality of all $Q^V_{\ell m}$ to expand the mean squared distance between $\tilde\rho_V$ and $\rho$ as
+\begin{align}
+  \E_V\p{\rho}^2 \equiv \Bbk{\norm{\tilde\rho_V-\rho}^2}
   = \sum_{\ell,m,i,i'} \p{q^V_{\ell mi}}^* q^V_{\ell mi'} \bbk{\epsilon_{i\ell} \epsilon_{i'\ell}}
-  = \f1n \sum_{\ell,i} \abs{\tilde{\v q}_{\ell i}}^2 \cov_\rho\p{T_{i\ell},T_{i\ell}},
-  &&
-  \abs{\tilde{\v q}_{\ell i}}^2 = \sum_m \abs{q_{\ell mi}}^2.
+  = \sum_{\ell,m,i} \abs{q^V_{\ell mi}}^2 \bbk{\epsilon_{i\ell}^2}
+  < \f1n \sum_\ell \Gamma_\ell^2 \S_{V\ell}^2,
+  \label{eq:bound_apndx}
+\end{align}
+where we used the fact that $\bbk{\epsilon_{i\ell}^2}\le\Gamma_\ell^2/n$, and
+\begin{align}
+  \sum_{m,i} \abs{q^V_{\ell mi}}^2
+  = \sum_m \f1{\p{M^V_{\ell m}}^2}
+  = \norm{M_{V\ell}^{-1}}
+  = \S_{V\ell}^2.
+\end{align}
+The fact that $\bbk{\epsilon_{i\ell}^2}<1/2n$ also implies that
+\begin{align}
+  \E_V\p{\rho}^2 < \f1{2n} \sum_\ell \S_{V\ell}^2 = \f{\S_V^2}{2n}.
+\end{align}
+Note that the bound in Eq.~\eqref{eq:bound_apndx} is not tight, as the individual bounds on the variances $\bbk{\epsilon_{i\ell}^2}$ cannot all be achieved simultaneously.
+There is therefore still room for improvement on the bound in Eq.~\eqref{eq:bound} by maximizing $\E_V$ over the set of all qudit states $\rho$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Exact reconstruction error}
+\label{sec:exact}
+
+Here we find exact expressions for reconstruction error, which can be used to estimate the error in a given reconstruction $\tilde\rho_V$ of an unknown state $\rho$ after performing tomography.
+To this end, we start with Eq.~\eqref{eq:bound_apndx} from Appendix \ref{sec:revisiting} to write
+\begin{align}
+  \E_V\p{\rho}^2
+  = \sum_{\ell,m,i,i'} \p{q^V_{\ell mi}}^* q^V_{\ell mi'} \bbk{\epsilon_{i\ell}\epsilon_{i'\ell}}
+  = \f1n \sum_{\ell,m,i} \abs{\tilde{\v q}_{\ell i}}^2 \cov_\rho\p{T_{i\ell},T_{i\ell}},
+  &&
+  \abs{\tilde{\v q}_{\ell i}}^2 = \sum_m \abs{q_{\ell mi}}^2,
   \label{eq:error_SM_start}
 \end{align}
-where $\tilde{\v q}_{\ell i} = (q_{\ell,-\ell,i},q_{\ell,-\ell+1,i},\cdots q_{\ell,\ell,i})$.
-Identifying the singular value decomposition $M_{V\ell} = U_{V\ell} \Sigma_{V\ell} W_{V\ell}^\dag$, the we observe that vector $\tilde{\v q}_{\ell i}$ is essentially the $i$-th row of $\Sigma_{V\ell}^{-1} W_{V\ell}^\dag$, which is to say that $\tilde{\v q}_{\ell i} = \Sigma_{V\ell}^{-1} W_{V\ell}^\dag \ket{\v v_i}$.
+where $\tilde{\v q}_{\ell i} = (q_{\ell,-\ell,i},q_{\ell,-\ell+1,i},\cdots q_{\ell,\ell,i})$, and we used the fact that $\bbk{\epsilon_{i\ell}\epsilon_{i'\ell}} = \delta_{ii'}\times\cov_\rho\p{T_{i\ell},T_{i\ell}}/n$.
+Identifying the singular value decomposition $M_{V\ell} = U_{V\ell} \Sigma_{V\ell} W_{V\ell}^\dag$, we then we observe that vector $\tilde{\v q}_{\ell i}$ is essentially the $i$-th row of $\Sigma_{V\ell}^{-1} W_{V\ell}^\dag$, which is to say that $\tilde{\v q}_{\ell i} = \Sigma_{V\ell}^{-1} W_{V\ell}^\dag \ket{\v v_i}$.
 We can therefore simplify
@@
 \begin{align}
-  \Bbk{\norm{\tilde\rho_V-\rho}^2}
+  \E_V\p{\rho}^2
   = \f1n \sum_{i,\ell,m,m'}
@@
 \end{align}
-Altogether, the exact reconstruction error in Eq.~\eqref{eq:error_SM_start} can be expressed more compactly by defining the covariance matrix
+Altogether, this reconstruction error can be expressed more compactly by defining the {\it covariance matrix}
 \begin{align}
@@
 \end{align}
-and the noise matrix
+and the {\it noise matrix}
 \begin{align}
@@
 \begin{align}
-  \Bbk{\norm{\tilde\rho_V-\rho}^2}
-  = \f1n \sum_\ell \obk{\N_{V\ell}|\C_\ell\sp{\rho}},
-  \label{eq:error_SM}
+  \E_V\p{\rho}^2 = \f1n \sum_\ell \obk{\N_{V\ell}|\C_\ell\sp{\rho}},
+  \label{eq:error_SM_mid}
 \end{align}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{An improved reconstruction error bound}
-
-We now seek to maximize the error in Eq.~\eqref{eq:error_SM} over all $\rho$, which will provide an upper bound on reconstruction error for a given set of axes $V$.
-To this end, we will first rewrite this error in a way that is more convenient to evaluate for any given $\rho$.
-Using the fact that
+The result in Eq.~\eqref{eq:error_SM_mid} essentially expresses reconstruction error as a weighted sum of the covariances $\cov_\rho\p{T_{\ell m},T_{\ell m'}}$, where the weights are given by the corresponding matrix elements of the noise matrix $\N_{V\ell}$.
+This expression is perhaps the most physically meaningful form of the reconstruction error $\E_V\p{\rho}^2$ that we will consider in this work, but in practice it turns out that Eq.~\eqref{eq:error_SM_mid} is inconvenient and inefficient to evaluate for any given state $\rho$.
+To find a more practical expression of reconstruction error, we use the fact that
 \begin{align}
@@
 \end{align}
-we can expand the covariance matrix as
+to expand the covariance matrix as
 \begin{align}
@@
 where we define the inversion operator $\I\equiv\sum_m \p{-1}^m\op{-m}{m}$.
-We can then expand the product $T_{\ell m} T_{\ell m'}$ to simplify
+We then expand the product $T_{\ell m'} T_{\ell m}$ as
 \begin{align}
@@
 where the (real) factors $f_{\ell m';\ell m}^{L,m'+m}$ are provided in Appendix \ref{sec:trans_prod}.
-Substituting the covariance matrix back into Eq.~\eqref{eq:error_SM} and replacing $\obk{T_{\ell m}|\rho}\to\rho_{\ell m}$, we get
-\begin{align}
-  n \Bbk{\norm{\tilde\rho_V-\rho}^2}
-  = \sum_{L,M} \p{\chi^V_{LM}}^* \rho_{LM}
+Substituting the covariance matrix back into Eq.~\eqref{eq:error_SM_mid} and replacing $\obk{T_{\ell m}|\rho}\to\rho_{\ell m}$, we get
+\begin{align}
+  \E_V\p{\rho}^2
+  = \f1n \sp{\sum_{\ell,m} \p{\chi^V_{\ell m}}^* \rho_{\ell m}
   - \sum_{\ell,m,m'} \bk{m'|\I\N_{V\ell}|m}
-  \rho_{\ell m} \rho_{\ell m'},
+  \rho_{\ell m} \rho_{\ell m'}},
   \label{eq:error_comps}
@@
 \end{align}
-is expressed in terms of the matrices
+can be written in terms of the matrices
 \begin{align}
@@
 \end{align}
-Here $\D_M$ simply picks off the $M$-th diagonal of the matrix it acts on.
+Here $\D_M$ simply picks off the $M$-th diagonal of the matrix it acts on, such that $\obk{\N_{V\ell}|\D_M|\I g_{L\ell}}$ is an inner product of the $M$-th diagonal of $\I g_{L\ell}$ with the $\p{-M}$-th diagonal of $\N_{V\ell}$.
 Defining the $\p{2\ell+1}$-component vectors
@@
 \end{align}
-the expansion in Eq.~\eqref{eq:error_comps} is most efficiently evaluated in the vectorized form
-\begin{align}
-  \Bbk{\norm{\tilde\rho_V-\rho}^2}
+we can write the expansion in Eq.~\eqref{eq:error_comps} in the vectorized form
+\begin{align}
+  \E_V\p{\rho}^2
   = \f1n \sum_\ell \sp{\bk{\chi_{V\ell}|\rho_\ell}
     - \bk{\rho_\ell|\N_{V\ell}|\rho_\ell}}.
-\end{align}
-This error is straightforward to maximize by considering its variation with respect to $\rho_{\ell m}$:
-\begin{align}
-  n\, \f{\partial}{\partial\rho_{\ell m}}
-  \Bbk{\norm{\tilde\rho_V-\rho}^2}
-  = \p{\chi^V_{\ell m}}^*
-  - \sum_{m'} \sp{\bk{m'|\I\N_{V\ell}|m} + \bk{m|\I\N_{V\ell}|m'}}
-  \rho_{\ell m'}.
-\end{align}
-The complex conjugate of this variation implies that the state $\rho_V^\star$ that maximizes reconstruction error satisfies
-\begin{align}
-  \ket{\chi_{V\ell}}
-  - \sp{\N_{V\ell} + \I N_{V\ell}^* \I} \ket{\rho_{V\ell}^\star}
-  = \ket{\chi_{V\ell}}
-  - 2 \N_{V\ell} \ket{\rho_{V\ell}^\star}
-  \stackrel{\ell>0}{=} 0,
-\end{align}
-which implies that
-\begin{align}
-  \ket{\rho_{V\ell}^\star} \stackrel{\ell>0}{=}
+  \label{eq:error_SM}
+\end{align}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Comments on a tight reconstruction error bound}
+
+In principle, maximizing the reconstruction error in Eq.~\eqref{eq:error_SM} over all qudit states $\rho$ would provide a tight upper bound on reconstruction error for any set of axes $V$.
+To simplify this task somewhat, we first maximize Eq.~\eqref{eq:error_SM} over all $\rho$ with $\tr\p\rho=1$: this maximum occurs at a ``state'' $\rho_V^\star$ whose components are defined by
+\begin{align}
+  \ket{\rho_{V\ell}^\star} \stackrel{\ell\ne0}{\equiv}
   \f12 \N_{V\ell}^{-1} \ket{\chi_{V\ell}},
   &&
-  \ket{\rho_{V,0}^\star} = \f1{\sqrt{d}} \ket{0},
-\end{align}
-where the $\ell=0$ component of $\rho_V^\star$ enforces $\tr\rho_V^\star=1$.
-We thus find that
-\begin{align}
-  \Bbk{\norm{\tilde\rho_V-\rho}^2}
-  \le \Bbk{\norm{\tilde\rho_V-\rho_V^\star}^2}
-  = \f{\epsilon_V^2}{n},
-  \label{eq:bound_SM}
-\end{align}
-where
-\begin{align}
-  \epsilon_V^2
-  \equiv \sum_{\ell>0} \f14 \bk{\chi_{V\ell}|\N_{V\ell}^{-1}|\chi_{V\ell}}
-  + \sp{\f{\bk{\chi_{V,0}|0}}{\sqrt{d}} - \f{\bk{0|\N_{V,0}|0}}{d}}
-  = \sum_{\ell>0} \sp{\f14 \bk{\chi_{V\ell}|\N_{V\ell}^{-1}|\chi_{V\ell}}
-    + \f1d \tr\p{\N_{V\ell}}}.
-\end{align}
-Note that the bound in Eq.~\eqref{eq:bound_SM} is obtained by maximizing reconstruction error over all possible $\rho$ (satisfying $\tr\rho=1$) regardless of their spectrum, which implies that the maximizer $\rho_V^\star$ may generally be an unphysical ``state'' with negative eigenvalues.
-In principle, there is therefore room for improvement on the bound in Eq.~\eqref{eq:bound_SM} by maximizing reconstruction error over the restricted domain of states with a non-negative spectrum.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Simpler (and looser) upper bounds}
-
-In order to find a simple bound on reconstruction error, we define the probability $p^i_\mu\equiv\bk{\Pi_{i\mu}}_\rho$, collect these probabilities into the classical probability distribution $p^i = \p{p^i_{-s}, p^i_{-s+1}, \cdots, p^i_s}$, and define the vector $t_\ell \equiv \p{t_{\ell,-s},t_{\ell,-s+1},\cdots,t_{\ell s}}$ of all $t_{\ell\mu}\equiv\bk{\mu|T_{\ell,0}|\mu}$.
-We then observe that
-\begin{align}
-  \bbk{\epsilon_{i\ell}^2} = \f1n \times \sigma_{p^i}^2\p{t_\ell},
-  &&
-  \sigma_p^2\p{X}
-  \equiv \sum_\mu p_\mu X_\mu^2 - \p{\sum_\mu p_\mu X_\mu}^2,
-\end{align}
-where $\sigma_p^2\p{X}$ is the weighted variance of $X$.
-This variance is maximal when $p$ has equal weight on the largest and smallest values of $X$, which implies that
-\begin{align}
-  \sigma_p^2\p{t_\ell} \le \Gamma_\ell^2,
-  &&
-  \Gamma_\ell \equiv \f{\max_\mu t_{\ell\mu} - \min_\mu t_{\ell\mu}}{2},
-\end{align}
-and lets us bound
-\begin{align}
-  \bbk{\epsilon_{i\ell}^2}
-  \le \f1n \times \Gamma_\ell^2
-  = \f{2\ell+1}{nd} \times \gamma_\ell^2,
-  \label{eq:variance_bound}
-\end{align}
-where $\gamma_\ell$ is defined using Clebsch-Gordan coefficients:
-\begin{align}
-  \gamma_\ell \equiv \f{\max_\mu\tau_{\ell\mu} - \min_\mu\tau_{\ell\mu}}{2},
-  &&
-  \tau_{\ell\mu} \equiv \bk{s\mu;\ell,0|s\mu}.
-\end{align}
-The bound in Eq.~\eqref{eq:variance_bound} lets us bound reconstruction error as
-\begin{align}
-  \Bbk{\norm{\tilde\rho_V-\rho}^2}
-  = \sum_{\ell,m,i} \abs{q^V_{\ell mi}}^2 \bbk{\epsilon_{i\ell}^2}
-  \le \f1n \sum_{\ell,m,i} \abs{q^V_{\ell mi}}^2 \Gamma_\ell^2
-  = \f1n \sum_\ell \Gamma_\ell^2 \S_{V\ell}^2
-  = \f1{nd} \sum_\ell \p{2\ell+1} \gamma_\ell^2 \S_{V\ell}^2,
-\end{align}
-where
-\begin{align}
-  \S_{V\ell}^2 = \sum_{m,i} \abs{q_{\ell mi}}^2
-  = \sum_m \f1{\p{M^V_{\ell m}}^2}
-  = \norm{M_{V\ell}^{-1}}^2.
-\end{align}
-To better understand this bound, we note that the Clebsch-Gordan coefficients $\tau_{\ell\mu}$ are at most $1$ in magnitude, which implies that $\gamma_\ell^2\le 1$.
-Moreover, we can also bound
-\begin{align}
-  \sigma_p^2\p{t_\ell} \le \sum_\mu p_\mu t_{\ell\mu}^2
-  \le \sum_\mu t_{\ell\mu}^2 = 1,
-  &&
-  \t{so}
-  &&
-  \Gamma_\ell^2 \le \min\Set{\f{2\ell+1}{d}, 1},
-  &&
-  \t{and}
-  &&
-  \Bbk{\norm{\tilde\rho_V-\rho}^2} \le \f{\S_V^2}{n}.
-\end{align}
-where $\S_V^2=\sum_\ell\S_{V\ell}^2=\norm{M_V^{-1}}^2$.
-We can get a tighter analytical bound on $\Gamma_\ell^2$ by considering the fact that $t_{\ell\mu}^2 = t_{\ell,-\mu}^2$ due to the symmetries of the Clebsch-Gordan coefficients.
-Denoting the index that maximizes $t_{\ell\mu}^2$ by $\mu_{\t{max}}$, it follows that if $\mu_{\t{max}}\ne0$ then
-\begin{align}
-  \sigma_p^2\p{t_\ell} \le \sum_\mu p_\mu t_{\ell\mu}^2
-  \le t_{\ell\mu_{\t{max}}}^2
-  = \f12 \p{t_{\ell\mu_{\t{max}}}^2 + t_{\ell,-\mu_{\t{max}}}^2}
-  \stackrel{\mu_{\t{max}}\ne0}{\le} \f12 \sum_\mu t_{\ell\mu}^2
-  = \f12.
-\end{align}
-If $\mu_{\t{max}}=0$, then similarly
-\begin{align}
-  t_{\ell\mu_{\t{max}}}^2 + 2 t_{\ell\mu_{\t{min}}}^2
-  = t_{\ell\mu_{\t{max}}}^2 + t_{\ell\mu_{\t{min}}}^2
-  + t_{\ell,-\mu_{\t{min}}}^2
-  \stackrel{\mu_{\t{max}}=0}{\le}
-  \sum_\mu t_{\ell\mu}^2 = 1,
-  &&
-  \t{so}
-  &&
-  \abs{t_{\ell\mu_{\t{min}}}}
-  \stackrel{\mu_{\t{max}}=0}{\le}
-  \sqrt{\f{1 - t_{\ell\mu_{\t{max}}}^2}{2}},
-\end{align}
-where $\mu_{\t{min}}$ is the index that minimizes $t_{\ell\mu}^2$.
-This inequality lets us bound
-\begin{align}
-  \Gamma_\ell
-  = \f12 \p{t_{\ell\mu_{\t{max}}} - t_{\ell\mu_{\t{min}}}}
-  \le \f12 \p{t_{\ell\mu_{\t{max}}} + \abs{t_{\ell\mu_{\t{min}}}}}
-  \stackrel{\mu_{\t{max}}=0}{\le}
-  \f12 t_{\ell\mu_{\t{max}}}
-  + \f12\sqrt{\f{1 - t_{\ell\mu_{\t{max}}}^2}{2}}
-  \equiv \lambda\p{t_{\ell\mu_{\t{max}}}}.
-\end{align}
-If is straightforward to show that $\lambda\p{x}$ has a single maximum at $x=\sqrt{2/3}$, and that $\lambda^\star\equiv \lambda\p{\sqrt{2/3}}=\sqrt{3/8}$, so
-\begin{align}
-  \Gamma_\ell^2
-  \stackrel{\mu_{\t{max}}=0}{\le} \p{\lambda^\star}^2
-  = \f38 < \f12.
-\end{align}
-Altogether, we thus find that in all cases
-\begin{align}
-  \Gamma_\ell^2 \le \min\Set{\f{2\ell+1}{d}, \f12},
-  &&
-  \t{so}
-  &&
-  \Bbk{\norm{\tilde\rho_V-\rho}^2}
-  \le \f{\S_V^2}{2n}.
-\end{align}
-The naive error bound in Eq.~\eqref{eq:bound_eps} of the main text therefore still holds with the replacement $\epsilon^2\to1/2n$.
+  \ket{\rho_{V,0}^\star} \equiv \f1{\sqrt{d}} \ket{0}.
+\end{align}
+The corresponding maximum of $\E_V$ is given by
+\begin{align}
+  \E_V\p{\rho_V^\star}^2 = \f1n \sum_{\ell>0}
+  \sp{\f14 \bk{\chi_{V\ell}|\N_{V\ell}^{-1}|\chi_{V\ell}}
+  - \f1d \tr\p{\N_{V\ell}}},
+\end{align}
+where the $\tr\p{\N_{V\ell}}$ terms above come from simplifying the $\ell=0$ terms of Eq.~\eqref{eq:error_SM} with $\rho\to\rho_V^\star$.
+While $\E_V\p{\rho_V^\star}^2$ provides an upper bound on $\E_V\p{\rho}^2$ over all $\rho$ with $\tr\p{\rho}=1$, this bound turns out to be useless in practice, because $\rho_V^\star$ will generally be a non-physical ``state'' with negative eigenvalues.
+To find tight bound on $\E_V\p{\rho}^2$ over the state of physical qudit states $\rho$, we also need to enforce the constraint that $\rho$ have no negative eigenvalues.
+Equipped with $\rho_V^\star$ and $\E_V\p{\rho_V^\star}^2$, we can expand
+\begin{align}
+  \E_V\p{\rho}^2 = \E_V\p{\rho_V^\star}^2
+  - \f1n D_V\p{\rho,\rho_V^\star}^2,
+  &&
+  D_V\p{X,Y}^2 \equiv
+  \sum_\ell \bk{\p{X-Y}_\ell | \N_{V\ell} | \p{X-Y}_\ell},
+\end{align}
+where $Z_\ell\equiv\sum_m \obk{T_{\ell m}|Z} \ket{m}$ is a vector of the degree-$\ell$ components of $Z$.
+Finding the maximum of $\E_V$ over all qudit states $\rho$ thus amounts to finding the closest physical qudit state $\rho$ to $\rho_V^\star$, with distance measured by the metric $D_V$ induced by the noise matrices $\N_{V\ell}$.
+This minimization problem is essentially solved for the case of $\N_{V\ell}=1$ in Ref.~\cite{smolin2012efficient}, but it is not obvious how to minimize $D_V\p{\rho,\rho_V^\star}^2$ for general $\N_{V\ell}$.
+We leave the general problem of minimizing $D_V\p{\rho,\rho_V^\star}$ to future work, noting that this task may be assisted by the positivity conditions derived in Ref.~\cite{kryszewski2006positivity}.
+A loose upper bound, which may nonetheless be tighter than that in Eq.~\eqref{eq:bound} of the main text, might be found by the fact that physical qudit states $\rho$ satisfy $\obk{\rho|\rho}\le1$.
 

### 2020-11-19 sun_phases/scripts/compute_error_scales.py

@@ -91,7 +91,7 @@ def compute_batch(pool, function, args):
 
 ##########################################################################################
 # pre-compute matrices of Wigner-3j coefficients,
-# which are used to build structure factor matrices
+# which are used to build matrices of structure constants
 
 def wigner_3j_mat(labels):
     LL, ll = labels
@@ -152,75 +152,95 @@ def rot_mid(LL, axis):
     return phases_1 * ( pulse_mats[LL] @ ( phases_0 * rot_zero_vecs[LL] ) )
 
 ##########################################################################################
-# compute the measurement matrix, its singular values, and the error scale
+# compute the measurement matrix, its singular values, and error scales
 
 # construct a fixed-degree measurement matrix
-def meas_mat(LL, axes):
+def degree_meas_mat(LL, axes):
     return np.array([ rot_mid(LL, axis) for axis in axes ])
 
 # get singular values ("norms") of a single fixed-degree measurement matrix
-def degree_norms(LL, axes):
-    return get_svdvals(meas_mat(LL, axes))
+def degree_meas_norms(LL, axes):
+    return get_svdvals(degree_meas_mat(LL, axes))
 
 # get singular values ("norms") of all fixed-degree measurement matrices
 pool = multiprocessing.Pool(processes = cpus)
 def meas_norms(dim, axes):
-    _degree_norms = functools.partial(degree_norms, axes = axes)
-    return compute_batch(pool, _degree_norms, range(dim-1,0,-1))
+    _degree_meas_norms = functools.partial(degree_meas_norms, axes = axes)
+    return compute_batch(pool, _degree_meas_norms, range(dim-1,-1,-1))
 
 # compute the classical error scale for a given set of axes
 def classical_error_scale(dim, axes):
-    norms = np.concatenate(list(meas_norms(dim, axes).values()))
-    return np.sqrt(sum(1/norms**2))
+    norms = meas_norms(dim, axes)
+    return np.sqrt(sum( sum(1/degree_norms**2)
+                        for degree_norms in norms.values() ))
+
+# squared "gamma" factors that appear in the quantum error scale
+def sqr_gamma(dim, ll):
+    ss = (dim-1)/2
+    mm_min, mm_max, wigner_3j_vals = wigner.lib.wigner_3j_m(ll, ss, ss, 0)
[…]

### 2020-11-19 sun_phases/scripts/multilevel_methods.py

@@ -7,6 +7,7 @@ from sympy.physics.wigner import wigner_6j as sym_wigner_6j
 
 # transition and drive operators, exact
 def transition_op_exact(dim, L, M):
+    if L >= dim or abs(M) > L: return np.zeros((dim,dim))
     L, M = sym.S(L), sym.S(M)
     I = sym.S(dim-1)/2
     mu_min = max(-M,0)
@@ -16,6 +17,7 @@ def transition_op_exact(dim, L, M):
     return sym.sqrt(sym.S(2*L+1)/sym.S(2*I+1)) * np.diag(diag_vals, -M)
 
 def drive_op_exact(dim, L, M):
+    if L >= dim or abs(M) > L: return np.zeros((dim,dim))
     T_LM = transition_op_exact(dim, L, M)
     if M == 0: return T_LM
     eta_M = (-1)**M if M > 0 else sym.I
@@ -43,7 +45,7 @@ def transition_prod_coef_exact(dim, l1, m1, l2, m2, L, M):
     I = sym.S(dim-1)/2
     return ( (-1)**(2*I+L) * sym.sqrt((2*l1+1)*(2*l2+1)) *
              sym_cg(l1, m1, l2, m2, L, M).doit() *
-             sym_wigner_6j(l1, l2, L, I, I, I).doit() )
+             sym_wigner_6j(l1, l2, L, I, I, I) )
 
 def transition_prod_coef(dim, l1, m1, l2, m2, L, M):
     return float(transition_prod_coef_exact(dim, l1, m1, l2, m2, L, M))

### 2020-11-20 sun_phases/scripts/compute_qudit_errors.py
rename some files/directories

@@ -30,7 +30,7 @@ try: seed = int(sys.argv[3])
 except: seed = 0
 np.random.seed(seed)
 
-data_dir = "../data/error_scale/"
+data_dir = "../data/qudit_errors/"
 if not os.path.isdir(data_dir):
     os.makedirs(data_dir)
 
@@ -233,7 +233,7 @@ def degree_chi_state(LL, noise_band_mats, inv_struct_bands):
     for ll, noise_band_mat in noise_band_mats.items():
         if ll < min_ll: continue
         _noise_band_mat = np.roll(noise_band_mat, -(2*ll+1)+LL, axis = 0)[:2*LL+1,:]
-        chi_state += (_noise_band_mat * inv_struct_bands[LL,ll]).sum(axis = 1)
+        chi_state += np.sum(_noise_band_mat * inv_struct_bands[LL,ll], axis = 1)
     return chi_state
 
 # compute the full "chi vector" in the "degree-order" basis

### 2020-11-20 sun_phases/scripts/plot_error_runtimes.py

@@ -4,8 +4,8 @@ import os, glob
 import numpy as np
 import matplotlib.pyplot as plt
 
-data_dir = "../data/error_scale/"
-fig_dir = "../figures/error_scale/"
+data_dir = "../data/qudit_errors/"
+fig_dir = "../figures/qudit_errors/"
 
 if not os.path.isdir(fig_dir):
     os.makedirs(fig_dir)

### 2020-11-20 sun_phases/spin_qudit_tomo.tex
rename some files/directories
renamed 'transition operators' to 'polarization operators' for consistency with existing literature

@@
 In this work, we provide an efficient framework for performing spin qudit tomography using only measurements of spin projection onto different spatial axes.
-We first introduce, in Section \ref{sec:transition_ops}, a basis of operators that have a simple physical interpretation, and transform nicely under rotations in 3D space, thereby respecting the symmetries of the tomography task at hand.
+We first introduce, in Section \ref{sec:polarization_ops}, a basis of operators that have a simple physical interpretation, and transform nicely under rotations in 3D space, thereby respecting the symmetries of the tomography task at hand.
 We then map this quantum tomography problem onto the well-studied classical problem of signal recovery on the sphere in Section \ref{sec:signal_recovery}, thereby importing a host of existing literature and mathematical machinery that has been developed for this task \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Spin transition operators and 3D rotations}
-\label{sec:transition_ops}
+\section{Polarization operators and 3D rotations}
+\label{sec:polarization_ops}
 
@@
 Due to the central importance of these transformation rules for a spin qudit, we first identify a basis of operators that transforms nicely under 3D rotations.
-One such basis is provided by the {\it transition operators} (also known as polarization operators \cite{kryszewski2006positivity, bertlmann2008bloch}), defined by
+One such basis is provided by the {\it polarization operators} \cite{kryszewski2006positivity, bertlmann2008bloch}, defined by
 \begin{align}
@@
 For brevity, we will generally treat the value of $d$ (or $s$) as constant but arbitrary throughout this work, and suppress any explicit dependence of quantities or operators such as $T_{\ell m}$ on $d$.
-The transition operators are orthonormal with respect to the trace inner product, and transform nicely under conjugation:
+The polarization operators are orthonormal with respect to the trace inner product, and transform nicely under conjugation:
 \begin{align}
@@
 where for any $d\times d$ matrix $X=\sum_{\mu,\nu} X_{\mu\nu} \op{\mu}{\nu}$ we define the $d^2$-component vector $\oket{X} \equiv \sum_{\mu,\nu} X_{\mu\nu} \ket{\mu\nu}$; $\obra{X}$ is the conjugate transpose of $\oket{X}$, such that $\obk{X|Y}=\tr\p{X^\dag Y}$; and $\delta_{kk'}\equiv 1$ if $k=k'$ and $0$ otherwise.
-These properties of the transition operators allow us to expand any density operator $\rho$ in the transition operator basis as
+These properties of the polarization operators allow us to expand any density operator $\rho$ in the polarization operator basis as
 \begin{align}
@@
 where $\bk{X}_\rho\equiv\tr\p{\rho X}=\obk{\rho|X}$, and $\rho^\dag=\rho$ implies that $\rho_{\ell m}^*=\p{-1}^m\rho_{\ell,-m}$.
-The transition operators can be physically interpreted as the transition $\ket\psi\to T_{\ell m}\ket\psi$ induced on the state $\ket\psi$ of a spin-$s$ system by the absorption of a spin-$\ell$ particle with spin projection $m$ onto a fixed quantization axis.
+The polarization operators can be physically interpreted as the polarization $\ket\psi\to T_{\ell m}\ket\psi$ induced on the state $\ket\psi$ of a spin-$s$ system by the absorption of a spin-$\ell$ particle with spin projection $m$ onto a fixed quantization axis.
 In analogy with the spherical harmonics $Y_{\ell m}$, we will refer to the index $\ell$ as the {\it degree} and $m$ as the {\it order} of $T_{\ell m}$.
 
-The transition operators are spherical tensor operators, whose degree is preserved under 3D rotations generated by the spin operators $S_\x,S_\y,S_\z$.
-Moreover, the degree-$\ell$ transition operators $T_{\ell m}$ transform identically to spin-$\ell$ particles under 3D rotations (see Appendix \ref{sec:rotations}).
+The polarization operators are spherical tensor operators, whose degree is preserved under 3D rotations generated by the spin operators $S_\x,S_\y,S_\z$.
+Moreover, the degree-$\ell$ polarization operators $T_{\ell m}$ transform identically to spin-$\ell$ particles under 3D rotations (see Appendix \ref{sec:rotations}).
 For any triplet of Euler angles $\v\omega=\p{\alpha_{\v\omega},\beta_{\v\omega},\gamma_{\v\omega}}$, we can therefore define the rotation operator
@@
 \end{align}
-and expand rotations of the transition operators as
+and expand rotations of the polarization operators as
 \begin{align}
@@
 For any fixed axis $\v v$, the sets $\set{\Pi_{\v v\mu}}$ and $\set{T_{\v v\ell,0}}$ are both orthonormal bases for the space of operators that are diagonal in the basis $\set{\ket{\mu_{\v v}}}$.
-By a simple change of operator basis, measuring the projectors $\set{\Pi_{\v v\mu}}$ is therefore equivalent to measuring the transition operators $\set{T_{\v v\ell,0}}$, and provides data on the expectation values $\bk{T_{\v v\ell,0}}_\rho$.
+By a simple change of operator basis, measuring the projectors $\set{\Pi_{\v v\mu}}$ is therefore equivalent to measuring the polarization operators $\set{T_{\v v\ell,0}}$, and provides data on the expectation values $\bk{T_{\v v\ell,0}}_\rho$.
 
@@
 \end{align}
-Expanding the rotated transition operators $T_{\v vk,0}$ into a sum of un-rotated transition operators, according to Eq.~\eqref{eq:trans_rot}, we find that the recovery condition in Eq.~\eqref{eq:state_recon} is satisfied when
+Expanding the rotated polarization operators $T_{\v vk,0}$ into a sum of un-rotated polarization operators, according to Eq.~\eqref{eq:trans_rot}, we find that the recovery condition in Eq.~\eqref{eq:state_recon} is satisfied when
 \begin{align}
@@
 \end{align}
-Orthogonality of the transition operators motivates the ansatz $C_{\ell mk}\p{\v v}=\delta_{k\ell}C_{\ell m}\p{\v v}$, which in turn implies that
+Orthogonality of the polarization operators motivates the ansatz $C_{\ell mk}\p{\v v}=\delta_{k\ell}C_{\ell m}\p{\v v}$, which in turn implies that
 \begin{align}
@@
 Moreover, for any fixed degree $\ell$, finding solutions to the reduced recovery problem in Eq.~\eqref{eq:tomo_recovery} is equivalent to the recovery of a degree-$\ell$ function $f = \sum_m f_m Y_{\ell m}$, which is provably possible with $\abs{V}=2\ell+1$ samples \cite{freeden2008spherical}.
-In the case of spin qudit tomography, the degree $\ell$ takes a maximal value of $\ell_{\t{max}}\equiv d-1$, so state recovery requires at least as many measurement axes as there are transition operators with degree $\ell_{\t{max}}$, namely $2\ell_{\t{max}}+1=2d-1$.
+In the case of spin qudit tomography, the degree $\ell$ takes a maximal value of $\ell_{\t{max}}\equiv d-1$, so state recovery requires at least as many measurement axes as there are polarization operators with degree $\ell_{\t{max}}$, namely $2\ell_{\t{max}}+1=2d-1$.
 
@@
 \begin{enumerate*}
-\item there are only $2\ell+1\le2d-1$ transition operators of degree $\ell\le\ell_{\t{max}}$, and
-\item measuring spin projection along any single axis provides measurement data for transition operators of {\it all} degrees $\ell\le\ell_{\t{max}}$,
+\item there are only $2\ell+1\le2d-1$ polarization operators of degree $\ell\le\ell_{\t{max}}$, and
+\item measuring spin projection along any single axis provides measurement data for polarization operators of {\it all} degrees $\ell\le\ell_{\t{max}}$,
 \end{enumerate*}
@@
 A set of measurement axes $V=\set{\v v}$ nominally induces a set of projectors $\set{\Pi_{\v v\mu}}$ that will be measured in an experiment.
-By a simple change of basis, measuring these projectors is equivalent to measuring the set of transition operators $\v T_V\equiv\set{T_{\v v\ell,0}}$.
+By a simple change of basis, measuring these projectors is equivalent to measuring the set of polarization operators $\v T_V\equiv\set{T_{\v v\ell,0}}$.
 By flattening the $d\times d$ matrices $T_{\v v\ell,0}$ in $\v T_V$ into $d^2$-component column vectors $\oket{T_{\v v\ell,0}}$, we can construct the {\it measurement matrix}
@@
 Here $\v v$ and $\ell$ label a column of $M_V$, or equivalently label a standard basis vector $\ket{\v v\ell}$ of a $\p{\abs{V}\times d}$-dimensional vector space.
-A necessary and sufficient condition for $V$ to allow for full state tomography is that the measured transition operators in $\v T_V$, or equivalently the columns of $M_V$, span the entire ($d^2$-dimensional) space of operators on a $d$-level spin qudit.
+A necessary and sufficient condition for $V$ to allow for full state tomography is that the measured polarization operators in $\v T_V$, or equivalently the columns of $M_V$, span the entire ($d^2$-dimensional) space of operators on a $d$-level spin qudit.
 In this case $M_V$ must be full rank, with $d^2$ nonzero singular values.
@@
 \end{align}
-where for shorthand we use a combined index $j=\p{\v v,\ell}$ to specify both a measurement axis $\v v$ and a degree $\ell$, which identify the transition operator $T_j\equiv T_{\v v\ell,0}$.
+where for shorthand we use a combined index $j=\p{\v v,\ell}$ to specify both a measurement axis $\v v$ and a degree $\ell$, which identify the polarization operator $T_j\equiv T_{\v v\ell,0}$.
 These operators allow us to expand any state $\rho$ of a $d$-level spin qudit in the form
@@
 Computing the classical error scale $\S_V^2$ to certify the bound in Eq.~\eqref{eq:bound_eps} requires constructing the measurement matrix $M_V$ and finding its singular values.
-The complexity of this task can be greatly reduced by the fact that the degree $\ell$ of a transition operator $T_{\ell m}$ is preserved under rotations, which implies that the unitary
+The complexity of this task can be greatly reduced by the fact that the degree $\ell$ of a polarization operator $T_{\ell m}$ is preserved under rotations, which implies that the unitary
 \begin{align}
@@
   \centering
-  \includegraphics{error_scale/error_times.pdf}
+  \includegraphics{qudit_errors/error_times.pdf}
   \caption{Serial runtime to compute the classical error scale $\S_V^2$ with $\abs{V}=2d-1$ randomly chosen measurement axes.
@@
 \end{align}
-Here $g_{L\ell}$ is a matrix of structure constants for the transition operator algebra (see Appendix \ref{sec:trans_prod}), and $\D_M$ simply picks off the $M$-th diagonal of the matrix it acts on, such that $\obk{\N_{V\ell}|\D_M|\I g_{L\ell}}$ is an inner product of the $M$-th diagonal of $\I g_{L\ell}$ with the $\p{-M}$-th diagonal of $\N_{V\ell}$.
+Here $g_{L\ell}$ is a matrix of structure constants for the polarization operator algebra (see Appendix \ref{sec:trans_prod}), and $\D_M$ simply picks off the $M$-th diagonal of the matrix it acts on, such that $\obk{\N_{V\ell}|\D_M|\I g_{L\ell}}$ is an inner product of the $M$-th diagonal of $\I g_{L\ell}$ with the $\p{-M}$-th diagonal of $\N_{V\ell}$.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Tomography protocols}
+\section{Tomography protocol}
 \label{sec:protocol}
 
-The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and optimize (using any standard minimization algorithm) the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the quantum error scale $\epsilon_V^2$ defined in Eq.~\eqref{eq:bound}.
+The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and optimize (using any standard minimization algorithm) the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the quantum error scale $\epsilon_V^2$.
 If $\abs{V}$ is too large for such optimization to be practical, one can instead generate many random sets of measurement axes, and then choose the set with the smallest quantum error scale $\epsilon_V^2$.
 After choosing a set of measurement axes $V$, (experimentally) make $n$ measurements of spin projection along every axis $\v v\in V$, and set $\tilde\Pi_{\v v\mu}\approx\bk{\Pi_{\v v\mu}}_\rho$ to the fraction of times in which the measurement outcome was $\mu$.
-The estimates $\tilde\Pi_{\v v\mu}$ of $\bk{\Pi_{\v v\mu}}_\rho$ determine estimates of the expectation values $\bk{T_{\v v\ell,0}}_\rho$, which can be used to construct an estimate $\tilde\rho_V$ of $\rho$, defined in Eq.~\eqref{eq:reconstructed_state}.
+The estimates $\tilde\Pi_{\v v\mu}$ of $\bk{\Pi_{\v v\mu}}_\rho$ determine estimates $\tilde T_{\v v\ell,0}$ of $\bk{T_{\v v\ell,0}}_\rho$, which can be used to construct the estimate $\tilde\rho_V$ of $\rho$
+The error in $\tilde\rho_V$ can be estimated by computing the right-hand side of Eq.~\eqref{eq:error} with the replacement $\rho_V\to\tilde\rho_V$.
 If $\tilde\rho_V$ has negative eigenvalues, its distance from $\rho$ can be reduced with maximum-likelihood corrections \cite{smolin2012efficient}.
-Finally, the error in $\tilde\rho_V$ can be estimated by computing the right-hand side of Eq.~\eqref{eq:error} with the replacement $\rho_V\to\tilde\rho_V$.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Rotating transition operators}
+\section{Rotating polarization operators}
 \label{sec:rotations}
@@
 
-Observing that $S_\z=T_{1,0}$ and $S_\pm\propto T_{1,\pm1}$, we can use the product operator expansion of the transition operators (see Appendix \ref{sec:trans_prod}), properties of Clebsch-Gordan coefficients, properties of Wigner $6$-$j$ symbols, and a computer algebra system to simplify the commutators
+Observing that $S_\z=T_{1,0}$ and $S_\pm\propto T_{1,\pm1}$, we can use the product operator expansion of the polarization operators (see Appendix \ref{sec:trans_prod}), properties of Clebsch-Gordan coefficients, properties of Wigner $6$-$j$ symbols, and a computer algebra system to simplify the commutators
 \begin{align}
@@
 which implies that $T_{\ell m}$ is a spherical tensor operator, whose degree degree $\ell$ is preserved under rotations generated by $\v S$.
-Moreover, by comparing Eqs.~\eqref{eq:spin_ops} and \eqref{eq:spin_trans} we see that the transition operators $T_{\ell m}$ transform identically to spin-$\ell$ particles under the adjoint action of the spin operators $S_\z$ and $S_\pm$.
+Moreover, by comparing Eqs.~\eqref{eq:spin_ops} and \eqref{eq:spin_trans} we see that the polarization operators $T_{\ell m}$ transform identically to spin-$\ell$ particles under the adjoint action of the spin operators $S_\z$ and $S_\pm$.
 For any triplet of Euler angles $\v\omega=\p{\alpha_{\v\omega},\beta_{\v\omega},\gamma_{\v\omega}}$, we can therefore define the fundamental rotation operator
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Variance of errors in the transition operator basis}
-
-Rather than the statistical errors $\epsilon_{i\mu} \equiv \tilde\Pi_{i\mu} - \bk{\Pi_{i\mu}}_\rho$ in the estimates $\tilde\Pi_{i\mu}$ of the projectors $\Pi_{i\mu}$, we now consider the statistical errors $\epsilon_{i\ell} \equiv \tilde T_{i\ell} - \bk{T_{i\ell}}_\rho$ in the estimates $\tilde T_{i\ell}$ of the transition operators $T_{i\ell} \equiv T_{\v v_i\ell,0}$.
-We can expand the transition operators $T_{i\ell}$ as a sum over projectors $\Pi_{i\mu}$ as
+\subsection{Variance of errors in the polarization operator basis}
+
+Rather than the statistical errors $\epsilon_{i\mu} \equiv \tilde\Pi_{i\mu} - \bk{\Pi_{i\mu}}_\rho$ in the estimates $\tilde\Pi_{i\mu}$ of the projectors $\Pi_{i\mu}$, we now consider the statistical errors $\epsilon_{i\ell} \equiv \tilde T_{i\ell} - \bk{T_{i\ell}}_\rho$ in the estimates $\tilde T_{i\ell}$ of the polarization operators $T_{i\ell} \equiv T_{\v v_i\ell,0}$.
+We can expand the polarization operators $T_{i\ell}$ as a sum over projectors $\Pi_{i\mu}$ as
 \begin{align}
@@
 and likewise $\tilde T_{i\ell} = \sum_\mu t_{\ell\mu} \tilde\Pi_{i\mu}$.
-The covariance between errors in the transition operator basis is then
+The covariance between errors in the polarization operator basis is then
 \begin{align}
@@
 where we used the fact that the covariance $\cov_\rho\p{X,Y}$ is linear in both $X$ and $Y$.
-Due to the appearance of $\delta_{ii'}$ above and the orthogonality of rotated transition operators $T_{i\ell}$ and $T_{i'\ell'}$ with degrees $\ell\ne\ell'$, only the variance $\bbk{\epsilon_{i\ell}^2}$ ultimately contributes to reconstruction error.
+Due to the appearance of $\delta_{ii'}$ above and the orthogonality of rotated polarization operators $T_{i\ell}$ and $T_{i'\ell'}$ with degrees $\ell\ne\ell'$, only the variance $\bbk{\epsilon_{i\ell}^2}$ ultimately contributes to reconstruction error.
 We therefore seek to find an upper bound on $\bbk{\epsilon_{i\ell}^2}$.
@@
 \end{align}
-To find a nicer analytical bound on $\bbk{\epsilon_{i\ell}^2}$, we can use normalization of the transition operators, $\obk{T_{i\ell}|T_{i\ell}}=\sum_\mu t_{\ell\mu}^2 = 1$, and the fact that all probabilities $p_\mu\le1$ to bound
+To find a nicer analytical bound on $\bbk{\epsilon_{i\ell}^2}$, we can use normalization of the polarization operators, $\obk{T_{i\ell}|T_{i\ell}}=\sum_\mu t_{\ell\mu}^2 = 1$, and the fact that all probabilities $p_\mu\le1$ to bound
 \begin{align}
@@
 where the $\tr\p{\N_{V\ell}}$ terms above come from simplifying the $\ell=0$ terms of Eq.~\eqref{eq:error_SM} with $\rho\to\rho_V^\star$.
-While $\E_V\p{\rho_V^\star}^2$ provides an upper bound on $\E_V\p{\rho}^2$ over all $\rho$ with $\tr\p{\rho}=1$, this bound turns out to be useless in practice, because $\rho_V^\star$ will generally be a non-physical ``state'' with negative eigenvalues.
-To find tight bound on $\E_V\p{\rho}^2$ over the state of physical qudit states $\rho$, we also need to enforce the constraint that $\rho$ have no negative eigenvalues.
+While $\E_V\p{\rho_V^\star}^2$ is a strict upper bound on $\E_V\p{\rho}^2$ over all $\rho$ with $\tr\p{\rho}=1$, this bound turns out to be useless in practice, because $\rho_V^\star$ will generally be a non-physical ``state'' with negative eigenvalues.
+To find tight bound on $\E_V\p{\rho}^2$ over the space of physical qudit states $\rho$, we also need to constrain $\rho$ to have no negative eigenvalues.
 Equipped with $\rho_V^\star$ and $\E_V\p{\rho_V^\star}^2$, we can expand
@@
 \end{align}
-where $Z_\ell\equiv\sum_m \obk{T_{\ell m}|Z} \ket{m}$ is a vector of the degree-$\ell$ components of $Z$.
+where $Z_\ell\equiv\sum_m \obk{T_{\ell m}|Z} \ket{m}$ is a vector of the degree-$\ell$ components of $Z$ in the polarization operator basis.
 Finding the maximum of $\E_V$ over all qudit states $\rho$ thus amounts to finding the closest physical qudit state $\rho$ to $\rho_V^\star$, with distance measured by the metric $D_V$ induced by the noise matrices $\N_{V\ell}$.
@@
 We leave the general problem of minimizing $D_V\p{\rho,\rho_V^\star}$ to future work, noting that this task may be assisted by the positivity conditions derived in Ref.~\cite{kryszewski2006positivity}.
-A loose upper bound, which may nonetheless be tighter than that in Eq.~\eqref{eq:bound} of the main text, might be found by the fact that physical qudit states $\rho$ satisfy $\obk{\rho|\rho}\le1$.
+A loose upper bound, which may nonetheless be tighter than that in Eq.~\eqref{eq:bound} of the main text (derived in Appendix \ref{sec:improved_bound}), might be found by the fact that physical qudit states $\rho$ satisfy $\obk{\rho|\rho}\le1$.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Transition operator product expansion}
+\section{Polarization operator product expansion}
 \label{sec:trans_prod}
 
-The transition operators on the $d$-dimensional Hilbert space of a spin-$s$ system (with $s\equiv\frac{d-1}{2}$) are defined by
+The polarization operators on the $d$-dimensional Hilbert space of a spin-$s$ system (with $s\equiv\frac{d-1}{2}$) are defined by
 \begin{align}
@@
 \end{align}
-we can find that the transition operators transform under conjugation as
+we can find that the polarization operators transform under conjugation as
 \begin{align}

### 2020-11-20 sun_phases/spin_qudit_tomo.tex
minor notation change

@@
 In principle, maximizing the reconstruction error in Eq.~\eqref{eq:error_SM} over all qudit states $\rho$ would provide a tight upper bound on reconstruction error for any set of axes $V$.
-To simplify this task somewhat, we first maximize Eq.~\eqref{eq:error_SM} over all $\rho$ with $\tr\p\rho=1$: this maximum occurs at a ``state'' $\rho_V^\star$ whose components are defined by
-\begin{align}
-  \ket{\rho_{V\ell}^\star} \stackrel{\ell\ne0}{\equiv}
+To simplify this task somewhat, we first maximize Eq.~\eqref{eq:error_SM} over all $\rho$ with $\tr\p\rho=1$: this maximum occurs at a ``state'' $\sigma_V^\star$ whose components are defined by
+\begin{align}
+  \ket{\sigma_{V\ell}^\star} \stackrel{\ell\ne0}{\equiv}
   \f12 \N_{V\ell}^{-1} \ket{\chi_{V\ell}},
   &&
-  \ket{\rho_{V,0}^\star} \equiv \f1{\sqrt{d}} \ket{0}.
+  \ket{\sigma_{V,0}^\star} \equiv \f1{\sqrt{d}} \ket{0}.
 \end{align}
@@
 \begin{align}
-  \E_V\p{\rho_V^\star}^2 = \f1n \sum_{\ell>0}
+  \E_V\p{\sigma_V^\star}^2 = \f1n \sum_{\ell>0}
   \sp{\f14 \bk{\chi_{V\ell}|\N_{V\ell}^{-1}|\chi_{V\ell}}
@@
 \end{align}
-where the $\tr\p{\N_{V\ell}}$ terms above come from simplifying the $\ell=0$ terms of Eq.~\eqref{eq:error_SM} with $\rho\to\rho_V^\star$.
-While $\E_V\p{\rho_V^\star}^2$ is a strict upper bound on $\E_V\p{\rho}^2$ over all $\rho$ with $\tr\p{\rho}=1$, this bound turns out to be useless in practice, because $\rho_V^\star$ will generally be a non-physical ``state'' with negative eigenvalues.
+where the $\tr\p{\N_{V\ell}}$ terms above come from simplifying the $\ell=0$ terms of Eq.~\eqref{eq:error_SM} with $\rho\to\sigma_V^\star$.
+While $\E_V\p{\sigma_V^\star}^2$ is a strict upper bound on $\E_V\p{\rho}^2$ over all $\rho$ with $\tr\p{\rho}=1$, this bound turns out to be useless in practice, because $\sigma_V^\star$ will generally be a non-physical ``state'' with negative eigenvalues.
 To find tight bound on $\E_V\p{\rho}^2$ over the space of physical qudit states $\rho$, we also need to constrain $\rho$ to have no negative eigenvalues.
-Equipped with $\rho_V^\star$ and $\E_V\p{\rho_V^\star}^2$, we can expand
-\begin{align}
-  \E_V\p{\rho}^2 = \E_V\p{\rho_V^\star}^2
-  - \f1n D_V\p{\rho,\rho_V^\star}^2,
+Equipped with $\sigma_V^\star$ and $\E_V\p{\sigma_V^\star}^2$, we can expand
+\begin{align}
+  \E_V\p{\rho}^2 = \E_V\p{\sigma_V^\star}^2
+  - \f1n D_V\p{\rho,\sigma_V^\star}^2,
   &&
@@
 where $Z_\ell\equiv\sum_m \obk{T_{\ell m}|Z} \ket{m}$ is a vector of the degree-$\ell$ components of $Z$ in the polarization operator basis.
-Finding the maximum of $\E_V$ over all qudit states $\rho$ thus amounts to finding the closest physical qudit state $\rho$ to $\rho_V^\star$, with distance measured by the metric $D_V$ induced by the noise matrices $\N_{V\ell}$.
-This minimization problem is essentially solved for the case of $\N_{V\ell}=1$ in Ref.~\cite{smolin2012efficient}, but it is not obvious how to minimize $D_V\p{\rho,\rho_V^\star}^2$ for general $\N_{V\ell}$.
-We leave the general problem of minimizing $D_V\p{\rho,\rho_V^\star}$ to future work, noting that this task may be assisted by the positivity conditions derived in Ref.~\cite{kryszewski2006positivity}.
+Finding the maximum of $\E_V$ over all qudit states $\rho$ thus amounts to finding the closest physical qudit state $\rho$ to $\sigma_V^\star$, with distance measured by the metric $D_V$ induced by the noise matrices $\N_{V\ell}$.
+This minimization problem is essentially solved for the case of $\N_{V\ell}=1$ in Ref.~\cite{smolin2012efficient}, but it is not obvious how to minimize $D_V\p{\rho,\sigma_V^\star}^2$ for general $\N_{V\ell}$.
+We leave the general problem of minimizing $D_V\p{\rho,\sigma_V^\star}$ to future work, noting that this task may be assisted by the positivity conditions derived in Ref.~\cite{kryszewski2006positivity}.
 A loose upper bound, which may nonetheless be tighter than that in Eq.~\eqref{eq:bound} of the main text (derived in Appendix \ref{sec:improved_bound}), might be found by the fact that physical qudit states $\rho$ satisfy $\obk{\rho|\rho}\le1$.

### 2020-11-20 sun_phases/spin_qudit_tomo.tex
plot all bound/reconsruction times

@@
   \centering
-  \includegraphics{qudit_errors/error_times.pdf}
-  \caption{Serial runtime to compute the classical error scale $\S_V^2$ with $\abs{V}=2d-1$ randomly chosen measurement axes.
+  \includegraphics{qudit_errors/qudit_times.pdf}
+  \caption{Serial runtime to compute $\S_V$, $\epsilon_V$, and $\E_V\p{\rho}$ with $\abs{V}=2d-1$ randomly chosen measurement axes and a randomly chosen qudit state $\rho$.
     Each point is an average over 100 calculations or 5 minutes of runtime, whichever comes first.
-    Dashed lines shows a fit to runtime $t=c d^\alpha$ for all $d\ge100$, finding $\alpha\approx 3.7$.
-    These results do not include fixed $O(d^3)$ runtimes for pre-computing quantities that can be recycled for every new choice of measurement axes $V$.}
+    Dashed lines shows a fit to runtime $t=c d^\alpha$ for all $d\ge100$, finding $\alpha\approx 3.7$ [\red{todo: update/correct after more simulations}].
+    These results do not count fixed $O(d^3)$ runtimes for pre-computing quantities that can be recycled for every new choice of $V$ or $\rho$.}
   \label{fig:times}
@@
 For comparison with previous error bound in Eq.~\eqref{eq:bound_eps}, we note that $\epsilon_V^2<\S_V^2/2$, so the previous bound still holds with the replacement $\epsilon^2\to1/2n$.
-The factors $\Gamma_\ell^2$ are quick to compute and can be recycled for every new choice of axes $V$, so the complexity of computing $\epsilon_V^2$ is identical to that of $\S_V^2$.
+The factors $\Gamma_\ell^2$ are quick to compute and can be recycled for every new choice of axes $V$, so the complexity of computing $\epsilon_V^2$ is identical to that of $\S_V^2$ (see Figure \ref{fig:times}).
 
@@
 Here $g_{L\ell}$ is a matrix of structure constants for the polarization operator algebra (see Appendix \ref{sec:trans_prod}), and $\D_M$ simply picks off the $M$-th diagonal of the matrix it acts on, such that $\obk{\N_{V\ell}|\D_M|\I g_{L\ell}}$ is an inner product of the $M$-th diagonal of $\I g_{L\ell}$ with the $\p{-M}$-th diagonal of $\N_{V\ell}$.
+Overall, the complexity of computing $\E_V\p{\rho}$ is identical to that of $\epsilon_V^2$ and $\S_V^2$ (see Figure \ref{fig:times}).
 

### 2020-11-20 sun_phases/scripts/plot_error_runtimes.py

@@ -15,35 +15,36 @@ params = { "font.size" : 10,
 plt.rcParams.update(params)
 
 figsize = (3,2)
-fit_dim_min = 100
 
 ##################################################
-
-files = glob.glob(data_dir + "times_*_serial.txt")
 
 def max_dim(file):
     pieces = file.replace(".txt","").split("/")[-1].split("_")
     return pieces[-1].split("-")[-1]
-files = sorted(files, key = max_dim)
-
-data = np.vstack([ np.loadtxt(file) for file in files ])
-sizes = data[:,0]
-times = data[:,1]
+def get_data(tag):
+    files = sorted(glob.glob(data_dir + f"times_{tag}_d*.txt"), key = max_dim)
+    data = np.vstack([ np.loadtxt(file) for file in files ])
+    return data[:,0], data[:,1]
 
 plt.figure(figsize = figsize)
-plt.loglog(sizes, times, "k.", label = "data")
 
-size_lims = plt.gca().get_xlim()
-time_lims = plt.gca().get_ylim()
+for tag, label, marker in [ ( "CB", r"$\mathcal{S}_V$", "ro" ),
+                            ( "QB", r"$\epsilon_V$", "k." ),
+                            ( "RE", r"$\mathcal{E}_V(\rho)$", "b." ) ]:
+    sizes, times = get_data(tag)
+    plot = plt.loglog(sizes, times, marker, label = label)
+    color = plot[0].get_color()
 
-fit_idx = sizes >= fit_dim_min
-fit = np.polyfit(np.log(sizes[fit_idx]), np.log(times[fit_idx]), deg = 1)
-plt.loglog(size_lims, np.exp(fit[0]*np.log(size_lims) + fit[1]), "k--", label = "fit")
-plt.gca().set_xlim(size_lims)
-plt.gca().set_ylim(time_lims)
+    size_lims = plt.gca().get_xlim()
+    time_lims = plt.gca().get_ylim()
+    fit_idx = sizes >= sizes[-1]//2
+    fit = np.polyfit(np.log(sizes[fit_idx]), np.log(times[fit_idx]), deg = 1)
+    plt.loglog(size_lims, np.exp(fit[0]*np.log(size_lims) + fit[1]), "--", color = color)
+    plt.gca().set_xlim(size_lims)
+    plt.gca().set_ylim(time_lims)
 
-plt.xlabel(r"qudit dimension $d$")
[…]

### 2020-11-20 sun_phases/spin_qudit_tomo.tex
add comment about Weyl-Wigner-Moyal formalism minor changes in notation

@@
   By mapping this quantum tomography problem onto the classical problem of signal recovery on the sphere, we prove that full reconstruction of arbitrary mixed states is possible, and requires a minimum number of measurement axes, $r_d^{\t{min}}$, that is bounded by $2d-1\le r_d^{\t{min}}\le d^2$.
-  We conjecture that $r_d^{\t{min}}=2d-1$, which we verify for all $d\le200$ using numerical experiments.
+  We conjecture that $r_d^{\t{min}}=2d-1$, which we verify for all $d\le200$ with numerical experiments.
   We then provide an algorithm with $O\p{rd^2}$ runtime for certifying an upper bound on the error with which measurements of spin projection along $r$ axes can reconstruct an unknown quantum state.
   Our algorithm motivates a simple randomized tomography protocol, which we use to study the trade-off between the number of measurement axes and state reconstruction error at a fixed total number of measurements [\red{note: this part is not yet complete}].
-  [\red{todo: maybe look into a simplified tomography procedure for pure states?]}
 \end{abstract}
@@
 Moreover, the degree-$\ell$ polarization operators $T_{\ell m}$ transform identically to spin-$\ell$ particles under 3D rotations (see Appendix \ref{sec:rotations}).
-For any triplet of Euler angles $\v\omega=\p{\alpha_{\v\omega},\beta_{\v\omega},\gamma_{\v\omega}}$, we can therefore define the rotation operator
-\begin{align}
-  R\p{\v\omega} \equiv e^{-\i\alpha_{\v\omega}S_\z} e^{-\i\beta_{\v\omega}S_\y} e^{-\i\gamma_{\v\omega}S_\z},
+For any triplet of Euler angles $\v\omega=\p{\alpha,\beta,\gamma}$, we can therefore define the rotation operator
+\begin{align}
+  R\p{\v\omega} \equiv e^{-\i\alpha S_\z} e^{-\i\beta S_\y} e^{-\i\gamma S_\z},
 \end{align}
@@
 are (Wigner) rotation matrix elements.
-Throughout this work, we will primarily consider rotations determined by points on the sphere.
-For ease of notation, we will therefore implicitly associate Euler angle doublets $\v v=\p{\alpha_{\v v},\beta_{\v v}}$ with the triplet $\p{\alpha_{\v v},\beta_{\v v},0}$, so that for example $R\p{\v v} = R\p{\alpha_{\v v},\beta_{\v v},0}$ and $T_{\v v\ell m} = T_{\p{\alpha_{\v v},\beta_{\v v},0},\ell m}$.
+As a final comment, we note that the polarization operators are closely related to the ``Weyl-Wigner-Moyal'' formalism for spin qudits \cite{li2013weylwignermoyal}, which provides spherical tensor operators with a phase-space representation that is essentially equal to the spherical harmonics $Y_{\ell m}$ (up to proportionality factors determined by $\ell$ and $m$).
+The precise connection between this formalism and polarization operators would be interesting to clarify in future work.
+
+Throughout this work, we will primarily consider rotations of the sphere that take the north pole to a fixed point $\v v=\p{\alpha,\beta}$ at azimuthal angle $\alpha$ and polar angle $\beta$.
+For ease of notation, we will therefore implicitly associate Euler angle doublets $\v v=\p{\alpha,\beta}$ with the triplet $\p{\alpha,\beta,0}$, such that $R\p{\v v} \equiv R\p{\alpha,\beta,0}$, $D^\ell_{mn}\p{\v v} \equiv D^\ell_{mn}\p{\alpha,\beta,0}$, and $T_{\v v\ell m} \equiv T_{\p{\alpha,\beta,0},\ell m}$.
+We will also primarily consider polarization operators with order $m=0$.
+Symbols with a suppressed order therefore implicitly set this order to zero, which is to say that $T_\ell\equiv T_{\ell,0}$, $T_{\v v\ell}\equiv T_{\v v\ell,0}$, and $D^\ell_n\equiv D^\ell_{n,0}$.
 
@@
 We are thus nominally restricted to measuring projectors $\Pi_{\v v\mu} \equiv \op{\mu_{\v v}}$, where $\ket{\mu_{\v v}}\equiv R\p{\v v}\ket{\mu}$ is a state with spin projection $\mu$ onto a measurement axis $\v v$ determined by a point on the sphere.
-For any fixed axis $\v v$, the sets $\set{\Pi_{\v v\mu}}$ and $\set{T_{\v v\ell,0}}$ are both orthonormal bases for the space of operators that are diagonal in the basis $\set{\ket{\mu_{\v v}}}$.
-By a simple change of operator basis, measuring the projectors $\set{\Pi_{\v v\mu}}$ is therefore equivalent to measuring the polarization operators $\set{T_{\v v\ell,0}}$, and provides data on the expectation values $\bk{T_{\v v\ell,0}}_\rho$.
+For any fixed axis $\v v$, the sets $\set{\Pi_{\v v\mu}}$ and $\set{T_{\v v\ell}}$ are both orthonormal bases for the space of operators that are diagonal in the basis $\set{\ket{\mu_{\v v}}}$.
+By a simple change of operator basis, measuring the projectors $\set{\Pi_{\v v\mu}}$ is therefore equivalent to measuring the polarization operators $\set{T_{\v v\ell}}$, and provides data on the expectation values $\bk{T_{\v v\ell}}_\rho$.
 
@@
 \begin{align}
-  \rho_{\ell m}^* = \bk{T_{\ell m}}_\rho
-  = \sum_{\v v,k} C_{\ell mk}\p{\v v} \bk{T_{\v vk,0}}_\rho.
+  \rho_{\ell m} = \bk{T_{\ell m}^\dag}_\rho
+  = \sum_{\v v,k} C_{\ell mk}\p{\v v} \bk{T_{\v vk}^\dag}_\rho.
   \label{eq:state_recon}
 \end{align}
-Expanding the rotated polarization operators $T_{\v vk,0}$ into a sum of un-rotated polarization operators, according to Eq.~\eqref{eq:trans_rot}, we find that the recovery condition in Eq.~\eqref{eq:state_recon} is satisfied when
-\begin{align}
-  T_{\ell m}
+Expanding the rotated polarization operators $T_{\v vk}$ into a sum of un-rotated polarization operators, according to Eq.~\eqref{eq:trans_rot}, we find that the recovery condition in Eq.~\eqref{eq:state_recon} is satisfied when
+\begin{align}
+  T_{\ell m}^\dag
   = \sum_{\v v,k,n} C_{\ell mk}\p{\v v}
-  D^k_{n,0}\p{\v v} T_{kn}.
+  D^k_n\p{\v v}^* T_{kn}^\dag.
 \end{align}
@@
 \begin{align}
-  \sum_{\v v} C_{\ell m}\p{\v v} D^\ell_{n,0}\p{\v v} = \delta_{mn}
+  \sum_{\v v} C_{\ell m}\p{\v v} D^\ell_n\p{\v v}^* = \delta_{mn}
   \label{eq:tomo_recovery}
@@
 We will refer to Eq.~\eqref{eq:full_recovery} as the {\it full recovery problem}, and Eq.~\eqref{eq:tomo_recovery} as the the {\it reduced recovery problem}.
-Due to the fact that $D^k_{n,0}\p{\v v} = \sqrt{\frac{4\pi}{2\ell+1}}\, Y_{kn}\p{\v v}^*$, any solution to the full recovery problem automatically solves the reduced recovery problem by setting $C_{\ell m}\p{\v v} = \sqrt{\frac{2\ell+1}{4\pi}}\, \tilde C_{\ell m}\p{\v v}^*$.
+Due to the fact that $D^k_n\p{\v v}^* = \sqrt{\frac{4\pi}{2\ell+1}}\, Y_{kn}\p{\v v}$, any solution to the full recovery problem automatically solves the reduced recovery problem by setting $C_{\ell m}\p{\v v} = \sqrt{\frac{2\ell+1}{4\pi}}\, \tilde C_{\ell m}\p{\v v}$.
 In principle, this mapping allows us to import a host of existing signal recovery algorithms \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality} for the task of spin qudit tomography.
@@
 Unfortunately, we have no general proof of this conjecture, which requires a deeper analysis of fixed-degree signal recovery on the sphere.
-Nonetheless, in the following section we numerically verify that $2d-1$ measurement axes suffice for all $d\le200$ [\red{todo: refer to Section \ref{sec:axes} instead}].
+Nonetheless, in the following sections we numerically verify that $2d-1$ measurement axes suffice for all $d\le200$.
 
@@
 A set of measurement axes $V=\set{\v v}$ nominally induces a set of projectors $\set{\Pi_{\v v\mu}}$ that will be measured in an experiment.
-By a simple change of basis, measuring these projectors is equivalent to measuring the set of polarization operators $\v T_V\equiv\set{T_{\v v\ell,0}}$.
-By flattening the $d\times d$ matrices $T_{\v v\ell,0}$ in $\v T_V$ into $d^2$-component column vectors $\oket{T_{\v v\ell,0}}$, we can construct the {\it measurement matrix}
-\begin{align}
-  M_V \equiv \sum_{\v v,\ell} \oket{T_{\v v\ell,0}} \bra{\v v\ell},
+By a simple change of basis, measuring these projectors is equivalent to measuring the set of polarization operators $\v T_V\equiv\set{T_{\v v\ell}}$.
+By flattening the $d\times d$ matrices $T_{\v v\ell}$ in $\v T_V$ into $d^2$-component column vectors $\oket{T_{\v v\ell}}$, we can construct the {\it measurement matrix}
+\begin{align}
+  M_V \equiv \sum_{\v v,\ell} \oket{T_{\v v\ell}} \bra{\v v\ell},
   \label{eq:meas_mat}
@@
 \end{align}
-where for shorthand we use a combined index $j=\p{\v v,\ell}$ to specify both a measurement axis $\v v$ and a degree $\ell$, which identify the polarization operator $T_j\equiv T_{\v v\ell,0}$.
+where for shorthand we use a combined index $j=\p{\v v,\ell}$ to specify both a measurement axis $\v v$ and a degree $\ell$, which identify the polarization operator $T_j\equiv T_{\v v\ell}$.
 These operators allow us to expand any state $\rho$ of a $d$-level spin qudit in the form
@@
 \end{align}
-where $\norm{X}^2 \equiv \obk{X|X} = \tr\p{X^\dag X}$ is the squared (Frobenius, or Hilbert-Schmidt) norm of $X$, and the {\it classical error scale}
+where $\norm{X}^2 \equiv \obk{X|X} = \tr\p{X^\dag X}$ is the squared (Frobenius, or Hilbert-Schmidt) norm of $X$, and the {\it classical error scale} $\S_V$ is defined by
 \begin{align}
@@
 is the squared norm of the right inverse $M_V^{-1}$ of $M_V$, satisfying $M_V M_V^{-1}=1$.
-We refer to this error scale $\S_V^2$ as ``classical'' because the bound in Eq.~\eqref{eq:bound_eps} applies in the presence of classical sources of measurement error.
-Note that the classical error scale $\S_V^2$ diverges if the measurement matrix $M_V$ is singular, in which case measuring spin projections along all axes in $V$ does not provide sufficient information to reconstruct arbitrary quantum states.
-
-Computing the classical error scale $\S_V^2$ to certify the bound in Eq.~\eqref{eq:bound_eps} requires constructing the measurement matrix $M_V$ and finding its singular values.
+We refer to the error scale $\S_V$ as ``classical'' because the bound in Eq.~\eqref{eq:bound_eps} applies in the presence of classical sources of measurement error.
+Note that the classical error scale $\S_V$ diverges if the measurement matrix $M_V$ is singular, in which case measuring spin projections along all axes in $V$ does not provide sufficient information to reconstruct arbitrary quantum states.
+
+Computing the classical error scale $\S_V$ to certify the bound in Eq.~\eqref{eq:bound_eps} requires constructing the measurement matrix $M_V$ and finding its singular values.
 The complexity of this task can be greatly reduced by the fact that the degree $\ell$ of a polarization operator $T_{\ell m}$ is preserved under rotations, which implies that the unitary
@@
 \begin{align}
-  M_{V\ell} \equiv \sum_{m,\v v} \ket{m} \obk{T_{\ell m}|T_{\v v\ell,0}} \bra{\v v}
+  M_{V\ell} \equiv \sum_{m,\v v} \ket{m} \obk{T_{\ell m}|T_{\v v\ell}} \bra{\v v}
   = \sum_{m,\v v} D^\ell_{m,0}\p{\v v} \op{m}{\v v}.
@@
 Constructing the block $M_{V\ell}$ and computing its singular values takes at most $O(\abs{V}d^2)$ time.
-If we assume that $\abs{V}\sim d$, then computing $\S_V^2$ to certify the bound in Eq.~\eqref{eq:bound_eps} takes $O(d^4)$ serial or $O(d^3)$ parallel runtime (see Figure \ref{fig:times}).
+If we assume that $\abs{V}\sim d$, then computing $\S_V$ to certify the bound in Eq.~\eqref{eq:bound_eps} takes $O(d^4)$ serial or $O(d^3)$ parallel runtime (see Figure \ref{fig:times}).
 
@@
   \includegraphics{qudit_errors/qudit_times.pdf}
-  \caption{Serial runtime to compute $\S_V$, $\epsilon_V$, and $\E_V\p{\rho}$ with $\abs{V}=2d-1$ randomly chosen measurement axes and a randomly chosen qudit state $\rho$.
+  \caption{Serial runtime to compute $\S_V$, $\epsilon_V$, or $\E_V\p{\rho}$ with $\abs{V}=2d-1$ randomly chosen measurement axes and a randomly chosen qudit state $\rho$.
     Each point is an average over 100 calculations or 5 minutes of runtime, whichever comes first.
     Dashed lines shows a fit to runtime $t=c d^\alpha$ for all $d\ge100$, finding $\alpha\approx 3.7$ [\red{todo: update/correct after more simulations}].
-    These results do not count fixed $O(d^3)$ runtimes for pre-computing quantities that can be recycled for every new choice of $V$ or $\rho$.}
+    These results do not count fixed $O(d^3)$ runtimes for pre-computing quantities that can be recycled for every new choice of $V$ or $\rho$.
+    The simulations used to benchmark runtime found axes $V$ with $\epsilon_V<d$ for all $d\le200$, thereby verifying for these cases that $2d-1$ measurement axes are sufficient to perform full state tomography.}
   \label{fig:times}
@@
 \end{align}
-where the {\it quantum error scale} $\epsilon_V^2$ is defined in terms of (half of) the spectral range of $T_{\ell,0}$:
+where the {\it quantum error scale} $\epsilon_V$ is defined in terms of (half of) the spectral range of $T_\ell$:
 \begin{align}
@@
   &&
-  t_{\ell\mu} \equiv \bk{\mu|T_{\ell,0}|\mu}.
+  t_{\ell\mu} \equiv \bk{\mu|T_\ell|\mu}.
 \end{align}
@@
 For comparison with previous error bound in Eq.~\eqref{eq:bound_eps}, we note that $\epsilon_V^2<\S_V^2/2$, so the previous bound still holds with the replacement $\epsilon^2\to1/2n$.
-The factors $\Gamma_\ell^2$ are quick to compute and can be recycled for every new choice of axes $V$, so the complexity of computing $\epsilon_V^2$ is identical to that of $\S_V^2$ (see Figure \ref{fig:times}).
-
-Though straightforward to compute, the bound in Eq.~\eqref{eq:bound} is not tight, as it is acquired by bounding the statistical variance $\epsilon_{\v v\ell}^2$ in the empirical estimate $\tilde T_{\v v\ell,0}$ of $\bk{T_{\v v\ell,0}}_\rho$ as $\bbk{\epsilon_{\v v\ell}^2}\le\Gamma_\ell^2$.
+The factors $\Gamma_\ell^2$ are quick to compute and can be recycled for every new choice of axes $V$, so the complexity of computing $\epsilon_V$ is the same as that of $\S_V$ (see Figure \ref{fig:times}).
+
+Though straightforward to compute, the bound in Eq.~\eqref{eq:bound} is not tight, as it is acquired by bounding the statistical variance $\epsilon_{\v v\ell}^2$ in the empirical estimate $\tilde T_{\v v\ell}$ of $\bk{T_{\v v\ell}}_\rho$ as $\bbk{\epsilon_{\v v\ell}^2}\le\Gamma_\ell^2$.
 Though the individual bounds on $\bbk{\epsilon_{\v v\ell}^2}$ for each axis $\v v$ and degree $\ell$ are tight, these bounds cannot all be achieved simultaneously.
@@
 
-The quantum error scale $\epsilon_V^2$ provides a pessimistic upper bound on statistical error, which can be calculated without prior knowledge of the true qudit state $\rho$.
+The quantum error scale $\epsilon_V$ provides a pessimistic upper bound on statistical error, which can be calculated without prior knowledge of the true qudit state $\rho$.
 The actual error in the reconstruction $\tilde\rho_V$ of a particular state $\rho$ may be considerably smaller.
@@
 where $\ket{\rho_\ell} \equiv \sum_m \rho_{\ell m}\ket{m}$, and $\ket{\chi_{V\ell}}$ and $\N_{V\ell}$ are defined below.
-While the true error in $\tilde\rho_V$ cannot be known exactly without knowing $\rho$, this error can be estimated by computing the right-hand side of Eq.~\eqref{eq:error} with the replacement $\rho\to\tilde\rho_V$.
+While the true error in $\tilde\rho_V$ cannot be known exactly without knowing $\rho$, this error can be estimated after constructing $\tilde\rho_V$ by computing $\E_V\p{\tilde\rho_V}$.
 
@@
 Here $g_{L\ell}$ is a matrix of structure constants for the polarization operator algebra (see Appendix \ref{sec:trans_prod}), and $\D_M$ simply picks off the $M$-th diagonal of the matrix it acts on, such that $\obk{\N_{V\ell}|\D_M|\I g_{L\ell}}$ is an inner product of the $M$-th diagonal of $\I g_{L\ell}$ with the $\p{-M}$-th diagonal of $\N_{V\ell}$.
-Overall, the complexity of computing $\E_V\p{\rho}$ is identical to that of $\epsilon_V^2$ and $\S_V^2$ (see Figure \ref{fig:times}).
+Overall, the complexity of computing $\E_V\p{\rho}$ is the same as that of $\epsilon_V$ and $\S_V$ (see Figure \ref{fig:times}).
 
@@
 
-The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and optimize (using any standard minimization algorithm) the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the quantum error scale $\epsilon_V^2$.
-If $\abs{V}$ is too large for such optimization to be practical, one can instead generate many random sets of measurement axes, and then choose the set with the smallest quantum error scale $\epsilon_V^2$.
+The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and optimize (using any standard minimization algorithm) the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the quantum error scale $\epsilon_V$.
+If $\abs{V}$ is too large for such optimization to be practical, one can instead generate many random sets of measurement axes, and then choose the set with the smallest quantum error scale $\epsilon_V$.
 After choosing a set of measurement axes $V$, (experimentally) make $n$ measurements of spin projection along every axis $\v v\in V$, and set $\tilde\Pi_{\v v\mu}\approx\bk{\Pi_{\v v\mu}}_\rho$ to the fraction of times in which the measurement outcome was $\mu$.
-The estimates $\tilde\Pi_{\v v\mu}$ of $\bk{\Pi_{\v v\mu}}_\rho$ determine estimates $\tilde T_{\v v\ell,0}$ of $\bk{T_{\v v\ell,0}}_\rho$, which can be used to construct the estimate $\tilde\rho_V$ of $\rho$
+The estimates $\tilde\Pi_{\v v\mu}$ of $\bk{\Pi_{\v v\mu}}_\rho$ determine estimates $\tilde T_{\v v\ell}$ of $\bk{T_{\v v\ell}}_\rho$, which can be used to construct the estimate $\tilde\rho_V$ of $\rho$
 The error in $\tilde\rho_V$ can be estimated by computing the right-hand side of Eq.~\eqref{eq:error} with the replacement $\rho_V\to\tilde\rho_V$.
@@
 
-As a last point, we mention that the quantum error scale $\epsilon_V^2$ can also be used to evaluate the ``quality'' of measurement axes $V$ chosen with existing methods for sparse signal recovery on the sphere \cite{rauhut2011sparse, alem2012sparse}.
+As a last point, we mention that the quantum error scale $\epsilon_V$ can also be used to evaluate the ``quality'' of measurement axes $V$ chosen with existing methods for sparse signal recovery on the sphere \cite{rauhut2011sparse, alem2012sparse}.
 In practice, we find that uniform random sampling outperforms these methods, in the sense that it generates low-error-scale sets of measurement axes with higher probability [\red{todo: add figure?}].
@@
 Even so, $2d-1$ measurement axes may not be the optimal number to use.
-Increasing the number of measurement axes generally decreases the quantum error scale $\epsilon_V^2$, but at the cost of having to estimate more observables.
+Increasing the number of measurement axes generally decreases the quantum error scale $\epsilon_V$, but at the cost of having to estimate more observables.
 At a fixed total number of measurements, increasing the number of measurement axes $\abs{V}$ reduces the number of measurements $n$ devoted to each axis $\v v\in V$.
@@
 
-The reconstruction error bound in Eq.~\eqref{eq:bound} nominally provides a straightforward answer: at a fixed total number of measurements, $N=n\abs{V}$, the number of measurement axes should be chosen to minimize $\epsilon_V^2/n\propto\epsilon_V^2\abs{V}$.
+The reconstruction error bound in Eq.~\eqref{eq:bound} nominally provides a straightforward answer: at a fixed total number of measurements, $N=n\abs{V}$, the number of measurement axes should be chosen to minimize $\E_V\p{\rho}^2<\epsilon_V^2/n\propto\epsilon_V^2\abs{V}$.
 
@@
 Moreover, by comparing Eqs.~\eqref{eq:spin_ops} and \eqref{eq:spin_trans} we see that the polarization operators $T_{\ell m}$ transform identically to spin-$\ell$ particles under the adjoint action of the spin operators $S_\z$ and $S_\pm$.
-For any triplet of Euler angles $\v\omega=\p{\alpha_{\v\omega},\beta_{\v\omega},\gamma_{\v\omega}}$, we can therefore define the fundamental rotation operator
-\begin{align}
-  R\p{\v\omega} \equiv e^{-\i\alpha_{\v\omega}S_\z} e^{-\i\beta_{\v\omega}S_\y} e^{-\i\gamma_{\v\omega}S_\z},
+For any triplet of Euler angles $\v\omega=\p{\alpha,\beta,\gamma}$, we can therefore define the fundamental rotation operator
+\begin{align}
+  R\p{\v\omega} \equiv
+  e^{-\i\alpha S_\z} e^{-\i\beta S_\y} e^{-\i\gamma S_\z},
 \end{align}
@@
 
-Rather than the statistical errors $\epsilon_{i\mu} \equiv \tilde\Pi_{i\mu} - \bk{\Pi_{i\mu}}_\rho$ in the estimates $\tilde\Pi_{i\mu}$ of the projectors $\Pi_{i\mu}$, we now consider the statistical errors $\epsilon_{i\ell} \equiv \tilde T_{i\ell} - \bk{T_{i\ell}}_\rho$ in the estimates $\tilde T_{i\ell}$ of the polarization operators $T_{i\ell} \equiv T_{\v v_i\ell,0}$.
+Rather than the statistical errors $\epsilon_{i\mu} \equiv \tilde\Pi_{i\mu} - \bk{\Pi_{i\mu}}_\rho$ in the estimates $\tilde\Pi_{i\mu}$ of the projectors $\Pi_{i\mu}$, we now consider the statistical errors $\epsilon_{i\ell} \equiv \tilde T_{i\ell} - \bk{T_{i\ell}}_\rho$ in the estimates $\tilde T_{i\ell}$ of the polarization operators $T_{i\ell} \equiv T_{\v v_i\ell}$.
 We can expand the polarization operators $T_{i\ell}$ as a sum over projectors $\Pi_{i\mu}$ as
@@
   &&
-  t_{\ell\mu} \equiv \bk{\mu|T_{\ell,0}|\mu}
-  = \sqrt{\f{2\ell+1}{d}} \bk{s\mu;\ell,0|s\mu},
+  t_{\ell\mu} \equiv \bk{\mu|T_\ell|\mu}
+  = \sqrt{\f{2\ell+1}{d}} \bk{s\mu;\ell|s\mu},
 \end{align}
@@
 
-To this end, we define the probability $p^i_\mu\equiv\bk{\Pi_{i\mu}}_\rho$, collect these probabilities into the classical probability distribution $p^i = \p{p^i_{-s}, p^i_{-s+1}, \cdots, p^i_s}$, and define the vector $t_\ell \equiv \p{t_{\ell,-s},t_{\ell,-s+1},\cdots,t_{\ell s}}$ of all $t_{\ell\mu}\equiv\bk{\mu|T_{\ell,0}|\mu}$.
+To this end, we define the probability $p^i_\mu\equiv\bk{\Pi_{i\mu}}_\rho$, collect these probabilities into the classical probability distribution $p^i = \p{p^i_{-s}, p^i_{-s+1}, \cdots, p^i_s}$, and define the vector $t_\ell \equiv \p{t_{\ell,-s},t_{\ell,-s+1},\cdots,t_{\ell s}}$ of all $t_{\ell\mu}\equiv\bk{\mu|T_\ell|\mu}$.
 We then observe that
@@
 \begin{align}
-  M_V \equiv \sum_{\v v,\ell} \oket{T_{\v v\ell,0}} \bra{\v v\ell},
+  M_V \equiv \sum_{\v v,\ell} \oket{T_{\v v\ell}} \bra{\v v\ell},
 \end{align}
@@
 The result in Eq.~\eqref{eq:error_SM_mid} essentially expresses reconstruction error as a weighted sum of the covariances $\cov_\rho\p{T_{\ell m},T_{\ell m'}}$, where the weights are given by the corresponding matrix elements of the noise matrix $\N_{V\ell}$.
-This expression is perhaps the most physically meaningful form of the reconstruction error $\E_V\p{\rho}^2$ that we will consider in this work, but in practice it turns out that Eq.~\eqref{eq:error_SM_mid} is inconvenient and inefficient to evaluate for any given state $\rho$.
+This expression is perhaps the most physically meaningful form of the reconstruction error $\E_V\p{\rho}$ that we will consider in this work, but in practice it turns out that Eq.~\eqref{eq:error_SM_mid} is inconvenient and inefficient to evaluate for any given state $\rho$.
 To find a more practical expression of reconstruction error, we use the fact that
@@
 where the $\tr\p{\N_{V\ell}}$ terms above come from simplifying the $\ell=0$ terms of Eq.~\eqref{eq:error_SM} with $\rho\to\sigma_V^\star$.
-While $\E_V\p{\sigma_V^\star}^2$ is a strict upper bound on $\E_V\p{\rho}^2$ over all $\rho$ with $\tr\p{\rho}=1$, this bound turns out to be useless in practice, because $\sigma_V^\star$ will generally be a non-physical ``state'' with negative eigenvalues.
-To find tight bound on $\E_V\p{\rho}^2$ over the space of physical qudit states $\rho$, we also need to constrain $\rho$ to have no negative eigenvalues.
-Equipped with $\sigma_V^\star$ and $\E_V\p{\sigma_V^\star}^2$, we can expand
+While $\E_V\p{\sigma_V^\star}$ is a strict upper bound on $\E_V\p{\rho}$ over all $\rho$ with $\tr\p{\rho}=1$, this bound turns out to be useless in practice, because $\sigma_V^\star$ will generally be a non-physical ``state'' with negative eigenvalues.
+To find tight bound on $\E_V\p{\rho}$ over the space of physical qudit states $\rho$, we also need to constrain $\rho$ to have no negative eigenvalues.
+Equipped with $\sigma_V^\star$ and $\E_V\p{\sigma_V^\star}$, we can expand
 \begin{align}
@@
 Finding the maximum of $\E_V$ over all qudit states $\rho$ thus amounts to finding the closest physical qudit state $\rho$ to $\sigma_V^\star$, with distance measured by the metric $D_V$ induced by the noise matrices $\N_{V\ell}$.
-This minimization problem is essentially solved for the case of $\N_{V\ell}=1$ in Ref.~\cite{smolin2012efficient}, but it is not obvious how to minimize $D_V\p{\rho,\sigma_V^\star}^2$ for general $\N_{V\ell}$.
-We leave the general problem of minimizing $D_V\p{\rho,\sigma_V^\star}$ to future work, noting that this task may be assisted by the positivity conditions derived in Ref.~\cite{kryszewski2006positivity}.
+This minimization problem is essentially solved for the case of $\N_{V\ell}=1$ in Ref.~\cite{smolin2012efficient}, but it is not obvious how to minimize $D_V\p{\rho,\sigma_V^\star}$ for general $\N_{V\ell}$.
+We leave the general problem of minimizing $D_V\p{\rho,\sigma_V^\star}$ to future work, noting that this task may be assisted by positivity conditions such as those derived in Ref.~\cite{kryszewski2006positivity}.
 A loose upper bound, which may nonetheless be tighter than that in Eq.~\eqref{eq:bound} of the main text (derived in Appendix \ref{sec:improved_bound}), might be found by the fact that physical qudit states $\rho$ satisfy $\obk{\rho|\rho}\le1$.

### 2020-11-20 sun_phases/scripts/compute_qudit_errors.py
export qudit tomography methods into a methods file

@@ -1,301 +1,0 @@
-#!/usr/bin/env python3
-
-import os, sys, time, functools
-import scipy, scipy.linalg
-import numpy as np
-
-import wigner.lib, py3nj
-
-# flags for what to compute:
-CB = "CB" # classical error bound
-QB = "QB" # quantum error bound
-RE = "RE" # reconstruction error
-
-compute = RE
-write_data = False
-
-sample_cap = 1000 # maximum number of times we choose a random set of measurement axes
-time_cap = 300 # maximum time to run per qubit dimension, in seconds
-
-min_dim = int(sys.argv[1])
-try: max_dim = int(sys.argv[2])
-except: max_dim = min_dim
-
-seed = 0 # random number seed. change at your own peril
-np.random.seed(seed)
-
-data_dir = "../data/qudit_errors/"
-if not os.path.isdir(data_dir):
-    os.makedirs(data_dir)
-
-data_file = data_dir + f"times_{compute}_d{min_dim}-{max_dim}.txt"
-
-genesis = time.time()
-
-##########################################################################################
-# general simulation methods
-
-# get squared singular values of a matrix
-def get_svdvals(matrix):
-    # return scipy.linalg.svdvals(matrix) # for some reason this is slower...
-    if matrix.shape[0] < matrix.shape[1]:
-        MM = matrix @ matrix.conj().T
-    else:
-        MM = matrix.conj().T @ matrix
-    return np.sqrt(scipy.linalg.eigvalsh(MM))
-
-# get diagonal bands of a matrix
-def diagonals(mat, band_min = None, band_max = None):
-    rows, cols = mat.shape
-    fill = np.zeros(((cols - 1), cols), dtype = mat.dtype)
-    stacked = np.vstack((mat, fill, mat))
-    major_stride, minor_stride = stacked.strides
-    strides = major_stride, minor_stride * (cols + 1)
-    shape = (rows + cols - 1, cols)
-    reversed_diags = np.lib.stride_tricks.as_strided(stacked, shape, strides)
-    diags = np.roll(np.flipud(reversed_diags), 1, axis = 0)
-    if band_min == None: band_min = -cols
-    if band_max == None: band_max = +rows
-    bands = band_max - band_min
-    return np.roll(diags, -band_min, axis = 0)[:bands+1,:]
-
-# generate random axes on the sphere by uniform sampling
[…]

### 2020-11-20 sun_phases/scripts/compute_tomo_times.py

@@ -1,0 +1,117 @@
+#!/usr/bin/env python3
+
+import os, sys, time
+import numpy as np
+
+genesis = time.time()
+
+# flags for what to compute:
+CB = "CB" # classical error bound
+QB = "QB" # quantum error bound
+RE = "RE" # reconstruction error
+
+compute = RE # default
+for idx, arg in enumerate(sys.argv):
+    if arg in [ CB, QB, RE ]:
+        compute = arg
+        sys.argv.remove(arg)
+
+save_data = False
+
+# get minimum / maximum qudit dimensions from command-line inputs
+min_dim = int(sys.argv[1])
+try: max_dim = int(sys.argv[2])
+except: max_dim = min_dim
+
+sample_cap = 1000 # maximum number of times we choose a random set of measurement axes
+time_cap = 300 # maximum time to run per qubit dimension, in seconds
+
+seed = 0 # set random number seed. change at your own peril
+np.random.seed(seed)
+
+# determine the directory and file for saving data
+data_dir = "../data/qudit_tomo/"
+if not os.path.isdir(data_dir):
+    os.makedirs(data_dir)
+data_file = data_dir + f"times_{compute}_d{min_dim}-{max_dim}.txt"
+
+##########################################################################################
+# import qudit tomography methods
+
+import qudit_tomo_methods as tomo
+
+random_axes = tomo.random_axes
+meas_mat, meas_scales = tomo.get_meas_methods(max_dim)
+
+if compute == CB:
+    error_scale = tomo.classical_error_scale
+if compute == QB:
+    error_scale = tomo.quantum_error_scale
+
+inv_struct_bands = tomo.get_struct_method(max_dim)
+random_state = tomo.random_state
+recon_error = tomo.recon_error
+
+##########################################################################################
+# simulate!
+
+if save_data:
+    # initialize data file
+    with open(data_file, "w") as file:
+        file.write(f"# sample_cap: {sample_cap}\n")
+        file.write(f"# time_cap: {time_cap} sec\n")
+        file.write(f"# seed: {seed}\n")
+        file.write(f"# dim, mean_time")
+        if compute in [ CB, QB ]:
+            file.write(f", error_scale")
+        file.write(f"\n")
+
+for dim in range(min_dim, max_dim+1):
+
+    if compute in [ CB, QB ]:
+        # initialize the smallest error scale we've seen
+        min_error_scale = np.inf
+
+    if compute == RE:
[…]

### 2020-11-20 sun_phases/scripts/qudit_tomo_methods.py

@@ -1,0 +1,229 @@
+#!/usr/bin/env python3
+
+import sys, time, functools
+import scipy, scipy.linalg
+import numpy as np
+
+import wigner.lib, py3nj
+
+##########################################################################################
+# general simulation methods
+
+# get squared singular values of a matrix
+def get_svdvals(matrix):
+    # return scipy.linalg.svdvals(matrix) # for some reason this is slower...
+    if matrix.shape[0] < matrix.shape[1]:
+        MM = matrix @ matrix.conj().T
+    else:
+        MM = matrix.conj().T @ matrix
+    return np.sqrt(scipy.linalg.eigvalsh(MM))
+
+# get diagonal bands of a matrix
+def diagonals(mat, band_min = None, band_max = None):
+    rows, cols = mat.shape
+    fill = np.zeros(((cols - 1), cols), dtype = mat.dtype)
+    stacked = np.vstack((mat, fill, mat))
+    major_stride, minor_stride = stacked.strides
+    strides = major_stride, minor_stride * (cols + 1)
+    shape = (rows + cols - 1, cols)
+    reversed_diags = np.lib.stride_tricks.as_strided(stacked, shape, strides)
+    diags = np.roll(np.flipud(reversed_diags), 1, axis = 0)
+    if band_min == None: band_min = -cols
+    if band_max == None: band_max = +rows
+    bands = band_max - band_min
+    return np.roll(diags, -band_min, axis = 0)[:bands+1,:]
+
+# generate random axes on the sphere by uniform sampling
+def random_axes(axis_num):
+    axes = np.random.rand(axis_num,2)
+    axes[:,0] = np.arccos(2*axes[:,0]-1) # polar angles
+    axes[:,1] *= 2*np.pi # azimuthal angles
+    return axes
+
+# generate a random (probably unphysical) qudit state,
+# represented by a point within a `dim^2`-dimensional hypersphere
+def random_state(dim):
+    point = np.array([ np.random.normal() for _ in range(dim**2) ])
+    point /= np.sqrt( np.random.exponential() + sum(point**2) )
+    # organize components into (2L+1)-sized vectors of "degree" L < dim
+    return { LL : point[ LL**2 : LL**2 + 2*LL+1 ] for LL in range(dim) }
+
+# run a batch of independent jobs, possibly in parallel (not yet implemented)
[…]

### 2020-11-21 sun_phases/spin_qudit_tomo.tex
export qudit tomography methods into a methods file
rearranged text about number of measurement axes

@@
 \documentclass[notitlepage,twocolumn]{revtex4-2}
-
-\usepackage{setspace} % to change text spacing
 
@@
 \usepackage{graphicx} % for figures
-\graphicspath{{./figures/}} % set path for all figures
+\graphicspath{{./figures/qudit_tomo/}} % set path for all figures
 
@@
 
-[\red{todo: investigate and comment on the connection between $T_{\ell m}$ and $Y_{\ell m}$?}]
-
 Consider a $d$-state spin qudit with total spin $s\equiv\frac{d-1}{2}$.
@@
 one might expect that $2d-1$ suitably chosen measurement axes would always suffice to perform full tomography of a $d$-level spin qudit.
-In the language of classical signal recovery, we would expect there to exist a fixed set of $2\ell_{\t{max}}+1$ points on the sphere that enable the recovery of {\it any} function with fixed degree $\ell\le\ell_{\t{max}}$ (given prior knowledge of $\ell$) \footnote{We could further conjecture that if {\it any} set of points $V\subset\Omega_2$ enables the recovery of degree-$\ell_{\t{max}}$ functions on the sphere, then $V$ enables the recovery degree-$\ell$ functions with $\ell\le\ell_{\t{max}}$.
+In the language of classical signal recovery, we would expect there to exist a set of $2\ell_{\t{max}}+1$ points on the sphere that enable the recovery of {\it any} function with fixed degree $\ell\le\ell_{\t{max}}$ (given prior knowledge of $\ell$) \footnote{We could further conjecture that if {\it any} set of points $V\subset\Omega_2$ enables the recovery of degree-$\ell_{\t{max}}$ functions on the sphere, then $V$ enables the recovery degree-$\ell$ functions with $\ell\le\ell_{\t{max}}$.
   However, this condition is stronger than necessary to guarantee full recovery of a $d$-level spin qudit state with $2d-1$ measurement axes, which only requires the existence of {\it some} such set of points.}.
 Unfortunately, we have no general proof of this conjecture, which requires a deeper analysis of fixed-degree signal recovery on the sphere.
-Nonetheless, in the following sections we numerically verify that $2d-1$ measurement axes suffice for all $d\le200$.
+Nonetheless, in the following sections we numerically verify that $2d-1$ points suffice for all $d\le200$.
 
@@
 A set of measurement axes $V=\set{\v v}$ nominally induces a set of projectors $\set{\Pi_{\v v\mu}}$ that will be measured in an experiment.
-By a simple change of basis, measuring these projectors is equivalent to measuring the set of polarization operators $\v T_V\equiv\set{T_{\v v\ell}}$.
-By flattening the $d\times d$ matrices $T_{\v v\ell}$ in $\v T_V$ into $d^2$-component column vectors $\oket{T_{\v v\ell}}$, we can construct the {\it measurement matrix}
+By a simple change of basis, measuring these projectors is equivalent to measuring the polarization operators $\set{T_{\v v\ell}}$.
+By flattening each $d\times d$ matrix $T_{\v v\ell}$ into the $d^2$-component column vector $\oket{T_{\v v\ell}}$, we can construct the {\it measurement matrix}
 \begin{align}
@@
 Here $\v v$ and $\ell$ label a column of $M_V$, or equivalently label a standard basis vector $\ket{\v v\ell}$ of a $\p{\abs{V}\times d}$-dimensional vector space.
-A necessary and sufficient condition for $V$ to allow for full state tomography is that the measured polarization operators in $\v T_V$, or equivalently the columns of $M_V$, span the entire ($d^2$-dimensional) space of operators on a $d$-level spin qudit.
+A necessary and sufficient condition for $V$ to allow for full state tomography is that the measured polarization operators in $\set{T_{\v v\ell}}$, or equivalently the columns of $M_V$, span the entire ($d^2$-dimensional) space of operators on a $d$-level spin qudit.
 In this case $M_V$ must be full rank, with $d^2$ nonzero singular values.
@@
 
-Computing the classical error scale $\S_V$ to certify the bound in Eq.~\eqref{eq:bound_eps} requires constructing the measurement matrix $M_V$ and finding its singular values.
+Computing the classical error scale $\S_V$ and finding estimates $\tilde\rho^V_k\approx\rho^V_k$ requires building the measurement matrix $M_V$ and computing its singular value decomposition.
 The complexity of this task can be greatly reduced by the fact that the degree $\ell$ of a polarization operator $T_{\ell m}$ is preserved under rotations, which implies that the unitary
@@
 where $M_{V\ell}^{-1}$ is the right inverse of $M_{V\ell}$.
-Constructing the block $M_{V\ell}$ and computing its singular values takes at most $O(\abs{V}d^2)$ time.
-If we assume that $\abs{V}\sim d$, then computing $\S_V$ to certify the bound in Eq.~\eqref{eq:bound_eps} takes $O(d^4)$ serial or $O(d^3)$ parallel runtime (see Figure \ref{fig:times}).
+Constructing the block $M_{V\ell}$ and computing its singular value decomposition takes at most $O(\abs{V}d^2)$ time.
+If we assume that $\abs{V}\sim d$, then computing the classical error scale $\S_V$ takes $O(d^4)$ serial or $O(d^3)$ parallel runtime (see Figure \ref{fig:times}).
 
@@
   \centering
-  \includegraphics{qudit_errors/qudit_times.pdf}
+  \includegraphics{qudit_times.pdf}
   \caption{Serial runtime to compute $\S_V$, $\epsilon_V$, or $\E_V\p{\rho}$ with $\abs{V}=2d-1$ randomly chosen measurement axes and a randomly chosen qudit state $\rho$.
@@
 \end{align}
-where the {\it quantum error scale} $\epsilon_V$ is defined in terms of (half of) the spectral range of $T_\ell$:
+where the {\it quantum error scale} $\epsilon_V$, so called because it is set by quantum noise, is defined in terms of (half of) the spectral range of $T_\ell$:
 \begin{align}
@@
 \end{align}
-If $d$ is even, then $\Gamma_\ell=\max_\mu t_{\ell\mu}$.
+If $d$ is even or $\ell$ is odd, then $\Gamma_\ell=\max_\mu t_{\ell\mu}$.
 For comparison with previous error bound in Eq.~\eqref{eq:bound_eps}, we note that $\epsilon_V^2<\S_V^2/2$, so the previous bound still holds with the replacement $\epsilon^2\to1/2n$.
@@
 After choosing a set of measurement axes $V$, (experimentally) make $n$ measurements of spin projection along every axis $\v v\in V$, and set $\tilde\Pi_{\v v\mu}\approx\bk{\Pi_{\v v\mu}}_\rho$ to the fraction of times in which the measurement outcome was $\mu$.
-The estimates $\tilde\Pi_{\v v\mu}$ of $\bk{\Pi_{\v v\mu}}_\rho$ determine estimates $\tilde T_{\v v\ell}$ of $\bk{T_{\v v\ell}}_\rho$, which can be used to construct the estimate $\tilde\rho_V$ of $\rho$
-The error in $\tilde\rho_V$ can be estimated by computing the right-hand side of Eq.~\eqref{eq:error} with the replacement $\rho_V\to\tilde\rho_V$.
+The estimates $\tilde\Pi_{\v v\mu}$ of $\bk{\Pi_{\v v\mu}}_\rho$ determine estimates $\tilde T_{\v v\ell} \equiv \sum_\mu \bk{\mu|T_\ell|\mu}\tilde\Pi_{\v v\mu}$ of $\bk{T_{\v v\ell}}_\rho$, which can in turn be used to construct the estimate $\tilde\rho_V$ of $\rho$.
+The total error in $\tilde\rho_V$, measured by its root-mean-square distance from $\rho$, is approximately $\E_V\p{\tilde\rho_V}\approx\E_V\p{\rho}$.
 If $\tilde\rho_V$ has negative eigenvalues, its distance from $\rho$ can be reduced with maximum-likelihood corrections \cite{smolin2012efficient}.
@@
 
-As a last point, we mention that the quantum error scale $\epsilon_V$ can also be used to evaluate the ``quality'' of measurement axes $V$ chosen with existing methods for sparse signal recovery on the sphere \cite{rauhut2011sparse, alem2012sparse}.
-In practice, we find that uniform random sampling outperforms these methods, in the sense that it generates low-error-scale sets of measurement axes with higher probability [\red{todo: add figure?}].
-Two observations may explain the poor performance of these sparse signal recovery methods.
-First, spin qudit tomography essentially boils down to recovering sparse signals with a particular harmonic structure, namely a fixed degree $\ell$.
-The existing sparse signal recovery algorithms do not assume or account for any harmonic structure, and may therefore perform worse than expected on this particular subset of signals.
-Second, spin qudit tomography requires recovering signals that are band-limited at degree $d$, but have $O(d)$ non-zero harmonic coefficients (out of $d^2$ coefficients in total).
-The existing sparse signal recovery methods, on the other hand, may only perform well for signals with sparsity much smaller than $d$.
-In any case, the fact that uniform sampling outperforms these methods only simplifies our randomized tomography protocol.
+The tomography protocol we outline above leaves open the question of {\it how many} measurement axes to choose.
+Though $2d-1$ measurement axes may be {\it sufficient} for full state tomography, this is not necessarily the {\it best} choice of $\abs{V}$.
+Increasing the number of measurement axes generally decreases the quantum error scale $\epsilon_V$, but at the cost of having to estimate more observables.
+At a fixed total number of measurements, increasing $\abs{V}$ reduces the number of measurements $n$ devoted to each axis $\v v\in V$.
+This trade-off begs the question: how should one choose the number of measurement axes, $\abs{V}$?
+
+The reconstruction error bound in Eq.~\eqref{eq:bound} nominally provides a straightforward answer: at a fixed total number of measurements, $N=n\abs{V}$, the number of measurement axes should be chosen to minimize the (squared) reconstruction error $\E_V\p{\rho}^2<\epsilon_V^2/n\propto\epsilon_V^2\abs{V}$.
+
+\begin{figure}
+  \centering
+  \includegraphics{qudit_axes.pdf}
+  \caption{Smallest ``normalized'' error scale $\beta(p)$ found with $p$ excess measurement axes, i.e.~with $\abs{V}=2d-1+p$.
+    Color indicates the qudit dimension $d$ as specified by the legend.
+    Here $\beta(p) \equiv \epsilon_{\t{min}}(p)\times\sqrt{2d-1+p}$ and $\epsilon_{\t{min}}(p) \equiv \min\set{\epsilon_V:\abs{V}=2d-1+p}$ is minimized over $10^3$ choices of random measurement axes or 5 minutes of runtime (for each $p$), whichever comes first.}
+  \label{fig:axes}
+\end{figure}
+
+\vspace{3cm}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{How many measurement axes?}
-\label{sec:axes}
-
-We conjecture, and numerically verified for $d\le200$, that measurements of spin projection along $2d-1$ axes suffice to reconstruct an arbitrary state of a $d$-level spin qudit.
-Even so, $2d-1$ measurement axes may not be the optimal number to use.
-Increasing the number of measurement axes generally decreases the quantum error scale $\epsilon_V$, but at the cost of having to estimate more observables.
-At a fixed total number of measurements, increasing the number of measurement axes $\abs{V}$ reduces the number of measurements $n$ devoted to each axis $\v v\in V$.
-This trade-off begs the question: how many measurement axes should be chosen to perform spin qudit tomography?
-
-The reconstruction error bound in Eq.~\eqref{eq:bound} nominally provides a straightforward answer: at a fixed total number of measurements, $N=n\abs{V}$, the number of measurement axes should be chosen to minimize $\E_V\p{\rho}^2<\epsilon_V^2/n\propto\epsilon_V^2\abs{V}$.
+\section{Conclusions}
+\label{sec:conclusions}
 

### 2020-11-22 sun_phases/spin_qudit_tomo.tex
finish (draft of) intro and the section on the tomography protocol

@@
 \newcommand{\I}{\mathcal{I}}
-\renewcommand{\L}{\mathcal{L}}
 \newcommand{\N}{\mathcal{N}}
@@
   We consider the task of performing quantum state tomography on a $d$-state spin qudit, using only measurements of spin projection onto different quantization axes.
-  By mapping this quantum tomography problem onto the classical problem of signal recovery on the sphere, we prove that full reconstruction of arbitrary mixed states is possible, and requires a minimum number of measurement axes, $r_d^{\t{min}}$, that is bounded by $2d-1\le r_d^{\t{min}}\le d^2$.
-  We conjecture that $r_d^{\t{min}}=2d-1$, which we verify for all $d\le200$ with numerical experiments.
-  We then provide an algorithm with $O\p{rd^2}$ runtime for certifying an upper bound on the error with which measurements of spin projection along $r$ axes can reconstruct an unknown quantum state.
-  Our algorithm motivates a simple randomized tomography protocol, which we use to study the trade-off between the number of measurement axes and state reconstruction error at a fixed total number of measurements [\red{note: this part is not yet complete}].
+  By an exact mapping onto the classical problem of signal recovery on the sphere, we prove that full reconstruction of arbitrary qudit states requires is possible with a minimal number of measurement axes, $r_d^{\t{min}}$, that is bounded by $2d-1\le r_d^{\t{min}}\le d^2$.
+  We conjecture that $r_d^{\t{min}}=2d-1$, which we verify numerically for all $d\le200$.
+  We then provide algorithms with $O\p{rd^3}$ serial runtime, parallelizable down to $O\p{rd^2}$, for (i) computing (a priori) upper bounds on the expected error with which spin projection measurements along $r$ given axes can reconstruct an unknown qudit state, and (ii) estimating (a posteriori) the statistical error in a reconstructed state.
+  Our algorithms motivate a simple randomized tomography protocol, for which we numerically find that using more measurement axes can yield substantial benefits that plateau after $r\approx3d$.
 \end{abstract}
@@
 The problem of qudit tomography in particular is not new, with an extensive literature on a variety of techniques \cite{manko1997spin, thew2002qudit, flammia2005minimal, salazar2012quantum, sosa-martinez2017quantum, ha2018minimal, evrard2019enhanced, stefano2019set, palici2020oam}.
-However, existing proposals either require the capability to perform essentially arbitrary operations on a qudit \cite{thew2002qudit, flammia2005minimal, salazar2012quantum, sosa-martinez2017quantum, ha2018minimal, stefano2019set, palici2020oam}, or rely on infinite-dimensional representations of a quantum spin \cite{manko1997spin, evrard2019enhanced}, resulting in tomographic protocols that can be highly inefficient or unachievable in practice.
-In particular, the protocols based on infinite-dimensional representations of a quantum spin require only the capability to measure spin projection onto different spatial axes, which is usually straightforward with a spin qudit.
-Nonetheless, these protocols obfuscate the basic requirements for performing full state tomography, provide no straightforward error bounds or guarantees of accuracy, and extract only a vanishing faction of the information contained in measurement data.
+However, existing protocols either require the capability to perform essentially arbitrary operations on a qudit \cite{thew2002qudit, flammia2005minimal, salazar2012quantum, sosa-martinez2017quantum, ha2018minimal, stefano2019set, palici2020oam}, or rely on infinite-dimensional representations of a quantum spin \cite{manko1997spin, evrard2019enhanced}, resulting in tomographic protocols that can be highly inefficient or unachievable in practice.
+The protocols based on infinite-dimensional representations of a quantum spin have the advantage of reconstructing its state from measurements of spin projection onto different spatial axes, which are generally accessible with any spin qudit.
+Nonetheless, these protocols obfuscate the actual requirements for performing full state tomography, provide no straightforward error bounds or guarantees of accuracy, and extract only a small faction of the information contained in measurement data.
 
 In this work, we provide an efficient framework for performing spin qudit tomography using only measurements of spin projection onto different spatial axes.
-We first introduce, in Section \ref{sec:polarization_ops}, a basis of operators that have a simple physical interpretation, and transform nicely under rotations in 3D space, thereby respecting the symmetries of the tomography task at hand.
-We then map this quantum tomography problem onto the well-studied classical problem of signal recovery on the sphere in Section \ref{sec:signal_recovery}, thereby importing a host of existing literature and mathematical machinery that has been developed for this task \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
+We first introduce, in Section \ref{sec:polarization_ops}, a set of qudit operators that have a simple physical interpretation, transform nicely under 3D rotations, and play a central role in our work.
+We then map the problem of spin qudit tomography onto the well-studied classical problem of signal recovery on the sphere in Section \ref{sec:signal_recovery}, thereby importing a host of existing literature and mathematical machinery that has been developed for this task \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
 In particular, this mapping allows us to bound the minimum number of measurement axes necessary to perform full tomography on a $d$-level spin qudit, $r_d^{\t{min}}$, by $2d-1\le r_d^{\t{min}}\le d^2$.
-
-[\red{finish intro}]
+In Section \ref{sec:error} we provide (a priori) upper bounds and (a posteriori) estimates of the statistical error in a qudit state reconstructed from measurements of spin projection along a given set of $r$ measurement axes.
+The capability to determine upper bounds on reconstruction error a priori motivates a simple randomized tomography protocol that we outline in Section \ref{sec:protocol}, and for which we numerically find that using more measurement axes yields substantial benefits that plateau after $r\approx3d$.
+
+% todo: comment on code availability
 
@@
 
-[\red{todo: add figure for signal recovery problem?}]
+% todo: add figure for signal recovery problem?
 
@@
 \section{Tomographic reconstruction error}
-\label{sec:bound}
+\label{sec:error}
 
@@
 However, this assumption breaks down when measurement error is limited by fundamental quantum (shot) noise.
-We relax the assumption of Eq.~\eqref{eq:error_assumption} in Appendix \ref{sec:improved_bound}, where we instead assume that $\tilde\rho_V$ is built from $n$ independent measurements of spin projection along every axis $\v v\in V$, with shot noise the dominant source of error.
+We relax the assumption of Eq.~\eqref{eq:error_assumption} in Appendix \ref{sec:bound}, where we instead assume that $\tilde\rho_V$ is built from $n$ independent measurements of spin projection along every axis $\v v\in V$, with shot noise the dominant source of error.
 In this case, the constraints that all $\bk{\Pi_{\v v\mu}}_\rho\ge0$ and $\tr\p\rho=1$ allow us to bound the mean squared distance between $\tilde\rho_V$ and $\rho$ as
@@
 \end{align}
-where the {\it quantum error scale} $\epsilon_V$, so called because it is set by quantum noise, is defined in terms of (half of) the spectral range of $T_\ell$:
+where the {\it quantum error scale} $\epsilon_V$, so called because it is set by quantum shot noise, is defined in terms of (half of) the spectral range of $T_\ell$:
 \begin{align}
@@
 The actual error in the reconstruction $\tilde\rho_V$ of a particular state $\rho$ may be considerably smaller.
-The exact mean squared distance between $\tilde\rho_V$ and $\rho$, derived in Appendix \ref{sec:exact}, can be written in the form
+The exact mean squared distance between $\tilde\rho_V$ and $\rho$, derived in Appendix \ref{sec:exact_error}, can be written in the form
 \begin{align}
@@
   \chi^V_{LM} &\equiv
-  \sum_\ell \obk{\N_{V\ell}|\D_M|\I g_{L\ell}}, \\
-  g_{L\ell} &\equiv \sum_{m,m'}
-  \obk{T_{L,m+m'}|T_{\ell m}T_{\ell m'}} \op{m}{m'}, \\
-  \I &\equiv \sum_m \p{-1}^m \op{-m}{m}, \\
+  \sum_\ell \obk{\N_{V\ell}|\D_M|\tilde g_{L\ell}}, \\
+  \tilde g_{L\ell} &\equiv \sum_{m,m'}
+  \obk{T_{L,m+m'}|T_{\ell m}^\dag T_{\ell m'}} \op{m}{m'}, \\
   \D_M &\equiv \sum_{m,m'} \delta_{M,m'-m} \op{mm'}.
 \end{align}
-Here $g_{L\ell}$ is a matrix of structure constants for the polarization operator algebra (see Appendix \ref{sec:trans_prod}), and $\D_M$ simply picks off the $M$-th diagonal of the matrix it acts on, such that $\obk{\N_{V\ell}|\D_M|\I g_{L\ell}}$ is an inner product of the $M$-th diagonal of $\I g_{L\ell}$ with the $\p{-M}$-th diagonal of $\N_{V\ell}$.
-Overall, the complexity of computing $\E_V\p{\rho}$ is the same as that of $\epsilon_V$ and $\S_V$ (see Figure \ref{fig:times}).
-
-[\red{todo: comment on code availability}]
+Here $\tilde g_{L\ell}$ is essentially a matrix of structure constants for the polarization operator algebra (see Appendix \ref{sec:trans_prod}), and $\D_M$ simply picks off the $M$-th diagonal of the matrix it acts on, such that $\obk{\N_{V\ell}|\D_M|\tilde g_{L\ell}}$ is an inner product of the $M$-th diagonal of $\tilde g_{L\ell}$ with the $\p{-M}$-th diagonal of $\N_{V\ell}$.
+Overall, the complexity of computing $\E_V\p{\rho}$ is the same as that of $\S_V$ and $\epsilon_V$ (see Figure \ref{fig:times}).
 
@@
 
-The tomography protocol we outline above leaves open the question of {\it how many} measurement axes to choose.
-Though $2d-1$ measurement axes may be {\it sufficient} for full state tomography, this is not necessarily the {\it best} choice of $\abs{V}$.
+The tomography protocol we outline above leaves open the question of {\it how many} measurement axes to use.
+Though $2d-1$ measurement axes may be sufficient to perform full state tomography, this is not necessarily the best choice of $\abs{V}$.
 Increasing the number of measurement axes generally decreases the quantum error scale $\epsilon_V$, but at the cost of having to estimate more observables.
@@
 
-The reconstruction error bound in Eq.~\eqref{eq:bound} nominally provides a straightforward answer: at a fixed total number of measurements, $N=n\abs{V}$, the number of measurement axes should be chosen to minimize the (squared) reconstruction error $\E_V\p{\rho}^2<\epsilon_V^2/n\propto\epsilon_V^2\abs{V}$.
-
 \begin{figure}
@@
   \includegraphics{qudit_axes.pdf}
-  \caption{Smallest ``normalized'' error scale $\beta(p)$ found with $p$ excess measurement axes, i.e.~with $\abs{V}=2d-1+p$.
-    Color indicates the qudit dimension $d$ as specified by the legend.
-    Here $\beta(p) \equiv \epsilon_{\t{min}}(p)\times\sqrt{2d-1+p}$ and $\epsilon_{\t{min}}(p) \equiv \min\set{\epsilon_V:\abs{V}=2d-1+p}$ is minimized over $10^3$ choices of random measurement axes or 5 minutes of runtime (for each $p$), whichever comes first.}
+  \caption{Empirical measurement-adjusted error scales $\tilde\beta(p)$ with $p$ excess measurement axes, determined by minimizing over $10^3$ choices of measurement axes $V$ or 5 minutes of runtime (for each $p$), whichever comes first.
+    Color indicates the qudit dimension $d$ in the legend.
+    The rapid initial drop in $\tilde\beta(p)$ implies that using more measurement axes can substantially lower the upper bound on reconstruction error provided in Eq.~\eqref{eq:bound}, and that these benefits plateau after $p\approx d$.}
   \label{fig:axes}
@@
 
-\vspace{3cm}
+The reconstruction error bound in Eq.~\eqref{eq:bound} nominally provides a straightforward answer: at a fixed total number of measurements, $N=n\abs{V}$, the number of measurement axes should be chosen to minimize the (squared) reconstruction error $\E_V\p{\rho}^2<\epsilon_V^2/n\propto\epsilon_V^2\abs{V}$.
+We therefore consider the measurement-adjusted error scale
+\begin{align}
+  \beta(p)
+  \equiv \min\set{\epsilon_V:\abs{V}=2d-1+p} \times \sqrt{2d-1+p}
+\end{align}
+as a function of $p$, the number of measurement axes exceeding $2d-1$.
+Though we cannot minimize over all suitable choices of measurement axes $V$ to compute $\beta(p)$, we can compute an empirical upper bound $\tilde\beta(p)\ge\beta(p)$ by minimizing over a large number of randomly chosen $V$.
+Figure \ref{fig:axes} shows the results of such empirical minimization, where we find that $\tilde\beta(p)$ drops substantially before plateauing at $p\approx d$, after which there are only marginal benefits to using more measurement axes.
+In the interest of reducing experimental complexity as well as the runtime of our tomography protocol, which grow linearly in $\abs{V}$, we therefore conclude that spin qudit tomography should be performed with $\abs{V}\approx3d$ measurement axes.
 
@@
 \section{Conclusions}
-\label{sec:conclusions}
 
@@
 \section{An improved reconstruction error bound}
-\label{sec:improved_bound}
-
-In Section \ref{sec:bound} of the main text, we provided a reconstruction error bound using the assumption of Eq.~\eqref{eq:error_assumption}, namely that expectation values derived from spin projection measurements can be estimated up to uncorrelated errors with maximal variance $\epsilon^2$.
+\label{sec:bound}
+
+In Section \ref{sec:error} of the main text, we provided a reconstruction error bound using the assumption of Eq.~\eqref{eq:error_assumption}, namely that expectation values derived from spin projection measurements can be estimated up to uncorrelated errors with maximal variance $\epsilon^2$.
 This assumption is reasonable if measurement error is dominated by experimental sources of noise, and it yields a simple derivation of the reconstruction bound in Eq.~\eqref{eq:bound_eps}.
@@
 \subsection{Revisiting the reconstruction error bound}
-\label{sec:revisiting}
-
-We now revisit the derivation of reconstruction error in Section \ref{sec:bound} to make use of the bounds on variances $\bbk{\epsilon_{i\ell}^2}$.
+\label{sec:bound}
+
+We now revisit the derivation of reconstruction error in Section \ref{sec:error} to make use of the bounds on variances $\bbk{\epsilon_{i\ell}^2}$.
 To recap, for a set of measurement axes $V=\set{\v v}$ and degrees $\ell\in\set{0,1,\cdots,d-1}$ we construct the measurement matrix
@@
 \section{Exact reconstruction error}
-\label{sec:exact}
+\label{sec:exact_error}
 
 Here we find exact expressions for reconstruction error, which can be used to estimate the error in a given reconstruction $\tilde\rho_V$ of an unknown state $\rho$ after performing tomography.
-To this end, we start with Eq.~\eqref{eq:bound_apndx} from Appendix \ref{sec:revisiting} to write
+To this end, we start with Eq.~\eqref{eq:bound_apndx} from Appendix \ref{sec:bound} to write
 \begin{align}
@@
 We leave the general problem of minimizing $D_V\p{\rho,\sigma_V^\star}$ to future work, noting that this task may be assisted by positivity conditions such as those derived in Ref.~\cite{kryszewski2006positivity}.
-A loose upper bound, which may nonetheless be tighter than that in Eq.~\eqref{eq:bound} of the main text (derived in Appendix \ref{sec:improved_bound}), might be found by the fact that physical qudit states $\rho$ satisfy $\obk{\rho|\rho}\le1$.
+A loose upper bound, which may nonetheless be tighter than that in Eq.~\eqref{eq:bound} of the main text (derived in Appendix \ref{sec:bound}), might be found by the fact that physical qudit states $\rho$ satisfy $\obk{\rho|\rho}\le1$.
 

### 2020-11-22 sun_phases/spin_qudit_tomo.tex
minor changes to the main text

@@
 \newcommand{\z}{\text{z}}
-\renewcommand{\d}{\text{d}}
 
@@
 
-\newcommand{\ZZ}{\mathbb{Z}}
-
 \def\obra#1{\mathinner{({#1}|}}
@@
 
-\usepackage{dsfont} % for identity operator
-\newcommand{\1}{\mathds{1}}
-
 %%% figures
@@
 \setlist[enumerate,1]{label={(\roman*)}} % default in-line numbering
-\setlist{nolistsep} % more compact spacing between environments
 
@@
 \begin{align}
-  \S_V^2 \equiv \sum_k\f1{\p{M^V_k}^2} = \norm{M_V^{-1}}^2,
-\end{align}
-is the squared norm of the right inverse $M_V^{-1}$ of $M_V$, satisfying $M_V M_V^{-1}=1$.
+  \S_V^2 \equiv \sum_k \p{M^V_k}^{-2} = \norm{M_V^{-1}}^2,
+\end{align}
+where $M_V^{-1}$ is the right inverse of $M_V$, satisfying $M_V M_V^{-1}=1$.
 We refer to the error scale $\S_V$ as ``classical'' because the bound in Eq.~\eqref{eq:bound_eps} applies in the presence of classical sources of measurement error.
-Note that the classical error scale $\S_V$ diverges if the measurement matrix $M_V$ is singular, in which case measuring spin projections along all axes in $V$ does not provide sufficient information to reconstruct arbitrary quantum states.
-
-Computing the classical error scale $\S_V$ and finding estimates $\tilde\rho^V_k\approx\rho^V_k$ requires building the measurement matrix $M_V$ and computing its singular value decomposition.
+Note that the classical error scale $\S_V$ diverges if the measurement matrix $M_V$ is singular, which indicates that measuring spin projections along all axes in $V$ does not provide sufficient information to reconstruct arbitrary quantum states.
+
+Computing the classical error scale $\S_V$ and estimates $\tilde\rho^V_k\approx\rho^V_k$ requires building the measurement matrix $M_V$ and computing its singular value decomposition.
 The complexity of this task can be greatly reduced by the fact that the degree $\ell$ of a polarization operator $T_{\ell m}$ is preserved under rotations, which implies that the unitary
@@
   M_{V\ell} \equiv \sum_{m,\v v} \ket{m} \obk{T_{\ell m}|T_{\v v\ell}} \bra{\v v}
-  = \sum_{m,\v v} D^\ell_{m,0}\p{\v v} \op{m}{\v v}.
+  = \sum_{m,\v v} D^\ell_m\p{\v v} \op{m}{\v v}.
 \end{align}
@@
   \caption{Serial runtime to compute $\S_V$, $\epsilon_V$, or $\E_V\p{\rho}$ with $\abs{V}=2d-1$ randomly chosen measurement axes and a randomly chosen qudit state $\rho$.
-    Each point is an average over 100 calculations or 5 minutes of runtime, whichever comes first.
-    Dashed lines shows a fit to runtime $t=c d^\alpha$ for all $d\ge100$, finding $\alpha\approx 3.7$ [\red{todo: update/correct after more simulations}].
-    These results do not count fixed $O(d^3)$ runtimes for pre-computing quantities that can be recycled for every new choice of $V$ or $\rho$.
-    The simulations used to benchmark runtime found axes $V$ with $\epsilon_V<d$ for all $d\le200$, thereby verifying for these cases that $2d-1$ measurement axes are sufficient to perform full state tomography.}
+    Each point is an average over $10^3$ calculations or 5 minutes of runtime, whichever comes first.
+    These results do not count fixed runtimes to pre-compute quantities that can be recycled for every new choice of $V$ or $\rho$.
+    Dashed lines show fits to a runtime $t=c d^\alpha$ for all $d\ge100$, finding $\alpha\approx 3.7$.
+    These simulations found axes $V$ with $\epsilon_V<d$ for all $d\le200$, thereby verifying for these cases that $2d-1$ measurement axes are sufficient to perform full state tomography.}
   \label{fig:times}
@@
 We relax the assumption of Eq.~\eqref{eq:error_assumption} in Appendix \ref{sec:bound}, where we instead assume that $\tilde\rho_V$ is built from $n$ independent measurements of spin projection along every axis $\v v\in V$, with shot noise the dominant source of error.
-In this case, the constraints that all $\bk{\Pi_{\v v\mu}}_\rho\ge0$ and $\tr\p\rho=1$ allow us to bound the mean squared distance between $\tilde\rho_V$ and $\rho$ as
+In this case, the constraints that $\tr\p\rho=1$ and $\bk{\Pi_{\v v\mu}}_\rho\ge0$ allow us to bound the mean squared distance between $\tilde\rho_V$ and $\rho$ as
 \begin{align}
@@
 If $d$ is even or $\ell$ is odd, then $\Gamma_\ell=\max_\mu t_{\ell\mu}$.
-For comparison with previous error bound in Eq.~\eqref{eq:bound_eps}, we note that $\epsilon_V^2<\S_V^2/2$, so the previous bound still holds with the replacement $\epsilon^2\to1/2n$.
+For comparison with the ``classical'' error bound in Eq.~\eqref{eq:bound_eps}, we note that $\epsilon_V^2<\S_V^2/2$, so the previous bound still holds with the replacement $\epsilon^2\to1/2n$.
 The factors $\Gamma_\ell^2$ are quick to compute and can be recycled for every new choice of axes $V$, so the complexity of computing $\epsilon_V$ is the same as that of $\S_V$ (see Figure \ref{fig:times}).
 
-Though straightforward to compute, the bound in Eq.~\eqref{eq:bound} is not tight, as it is acquired by bounding the statistical variance $\epsilon_{\v v\ell}^2$ in the empirical estimate $\tilde T_{\v v\ell}$ of $\bk{T_{\v v\ell}}_\rho$ as $\bbk{\epsilon_{\v v\ell}^2}\le\Gamma_\ell^2$.
+Though straightforward to compute, the bound in Eq.~\eqref{eq:bound} is not tight, as it is acquired by bounding the statistical error in the empirical estimate $\tilde T_{\v v\ell}$ of $\bk{T_{\v v\ell}}_\rho$ by $\bbk{\epsilon_{\v v\ell}^2}\le\Gamma_\ell^2$.
 Though the individual bounds on $\bbk{\epsilon_{\v v\ell}^2}$ for each axis $\v v$ and degree $\ell$ are tight, these bounds cannot all be achieved simultaneously.
@@
 The quantum error scale $\epsilon_V$ provides a pessimistic upper bound on statistical error, which can be calculated without prior knowledge of the true qudit state $\rho$.
-The actual error in the reconstruction $\tilde\rho_V$ of a particular state $\rho$ may be considerably smaller.
+The actual error in the reconstruction $\tilde\rho_V$ of a particular state $\rho$ may be considerably smaller, and depends on $\rho$ itself.
 The exact mean squared distance between $\tilde\rho_V$ and $\rho$, derived in Appendix \ref{sec:exact_error}, can be written in the form
@@
 where $\ket{\rho_\ell} \equiv \sum_m \rho_{\ell m}\ket{m}$, and $\ket{\chi_{V\ell}}$ and $\N_{V\ell}$ are defined below.
-While the true error in $\tilde\rho_V$ cannot be known exactly without knowing $\rho$, this error can be estimated after constructing $\tilde\rho_V$ by computing $\E_V\p{\tilde\rho_V}$.
+While the true error in $\tilde\rho_V$ cannot be known exactly without knowing $\rho$, this error can be estimated by $\E_V\p{\tilde\rho_V} \approx \E_V\p{\rho}$.
 
@@
 In the interest of reducing experimental complexity as well as the runtime of our tomography protocol, which grow linearly in $\abs{V}$, we therefore conclude that spin qudit tomography should be performed with $\abs{V}\approx3d$ measurement axes.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions}
 
@@
 \subsection{Revisiting the reconstruction error bound}
-\label{sec:bound}
+\label{sec:revisiting}
 
@@
   &&
-  M_{V\ell} = \sum_{m,\v v} D^\ell_{m,0}\p{\v v} \op{m}{\v v},
+  M_{V\ell} = \sum_{m,\v v} D^\ell_m\p{\v v} \op{m}{\v v},
 \end{align}
@@
   \sum_{m,i} \abs{q^V_{\ell mi}}^2
-  = \sum_m \f1{\p{M^V_{\ell m}}^2}
+  = \sum_m \p{M^V_{\ell m}}^{-2}
   = \norm{M_{V\ell}^{-1}}
@@
 Here we find exact expressions for reconstruction error, which can be used to estimate the error in a given reconstruction $\tilde\rho_V$ of an unknown state $\rho$ after performing tomography.
-To this end, we start with Eq.~\eqref{eq:bound_apndx} from Appendix \ref{sec:bound} to write
+To this end, we start with Eq.~\eqref{eq:bound_apndx} from Appendix \ref{sec:revisiting} to write
 \begin{align}
@@
   = \cov_\rho\p{T_{i\ell}^\dag,T_{i\ell}}
-  = \sum_{m,m'} D^\ell_{m,0}\p{\v v_i}^* D^\ell_{m',0}\p{\v v_i}
+  = \sum_{m,m'} D^\ell_m\p{\v v_i}^* D^\ell_{m'}\p{\v v_i}
   \cov_\rho\p{T_{\ell m}^\dag, T_{\ell m'}},
@@
   = \f1n \sum_{i,\ell,m,m'}
-  D^\ell_{m',0}\p{\v v_i} \abs{\tilde{\v q}_{\ell i}}^2
-  D^\ell_{m,0}\p{\v v_i}^*
+  D^\ell_{m'}\p{\v v_i} \abs{\tilde{\v q}_{\ell i}}^2
+  D^\ell_m\p{\v v_i}^*
   \cov_\rho\p{T_{\ell m}^\dag, T_{\ell m'}}.

### 2020-11-22 sun_phases/spin_qudit_tomo.tex
comment on the wigner function tomography work in schmied2011tomographic

@@
 
-The problem of qudit tomography in particular is not new, with an extensive literature on a variety of techniques \cite{manko1997spin, thew2002qudit, flammia2005minimal, salazar2012quantum, sosa-martinez2017quantum, ha2018minimal, evrard2019enhanced, stefano2019set, palici2020oam}.
-However, existing protocols either require the capability to perform essentially arbitrary operations on a qudit \cite{thew2002qudit, flammia2005minimal, salazar2012quantum, sosa-martinez2017quantum, ha2018minimal, stefano2019set, palici2020oam}, or rely on infinite-dimensional representations of a quantum spin \cite{manko1997spin, evrard2019enhanced}, resulting in tomographic protocols that can be highly inefficient or unachievable in practice.
+The problem of qudit tomography in particular is not new, with an extensive literature on a variety of techniques \cite{manko1997spin, schmied2011tomographic, evrard2019enhanced, flammia2005minimal, thew2002qudit, salazar2012quantum, sosa-martinez2017quantum, ha2018minimal, stefano2019set, palici2020oam}.
+However, existing protocols either rely infinite-dimensional representations of a quantum spin \cite{manko1997spin, schmied2011tomographic, evrard2019enhanced}, or require the capability to perform essentially arbitrary operations on a qudit \cite{thew2002qudit, flammia2005minimal, salazar2012quantum, sosa-martinez2017quantum, ha2018minimal, stefano2019set, palici2020oam}, resulting in tomographic protocols that can be highly inefficient or unachievable in practice.
 The protocols based on infinite-dimensional representations of a quantum spin have the advantage of reconstructing its state from measurements of spin projection onto different spatial axes, which are generally accessible with any spin qudit.
-Nonetheless, these protocols obfuscate the actual requirements for performing full state tomography, provide no straightforward error bounds or guarantees of accuracy, and extract only a small faction of the information contained in measurement data.
+Nonetheless, these protocols obfuscate the actual requirements for performing full state tomography, provide no straightforward error bounds or guarantees of accuracy, and (with the exception of Ref.~\cite{schmied2011tomographic}) extract only a small faction of the information contained in measurement data.
 

### 2020-11-22 sun_phases/spin_qudit_tomo.tex
changes to discussion about a tight reconstruction error bound

@@
   \E_V\p{\rho}^2 = \E_V\p{\sigma_V^\star}^2
-  - \f1n D_V\p{\rho,\sigma_V^\star}^2,
-  &&
-  D_V\p{X,Y}^2 \equiv
-  \sum_\ell \bk{\p{X-Y}_\ell | \N_{V\ell} | \p{X-Y}_\ell},
-\end{align}
-where $Z_\ell\equiv\sum_m \obk{T_{\ell m}|Z} \ket{m}$ is a vector of the degree-$\ell$ components of $Z$ in the polarization operator basis.
-Finding the maximum of $\E_V$ over all qudit states $\rho$ thus amounts to finding the closest physical qudit state $\rho$ to $\sigma_V^\star$, with distance measured by the metric $D_V$ induced by the noise matrices $\N_{V\ell}$.
-This minimization problem is essentially solved for the case of $\N_{V\ell}=1$ in Ref.~\cite{smolin2012efficient}, but it is not obvious how to minimize $D_V\p{\rho,\sigma_V^\star}$ for general $\N_{V\ell}$.
-We leave the general problem of minimizing $D_V\p{\rho,\sigma_V^\star}$ to future work, noting that this task may be assisted by positivity conditions such as those derived in Ref.~\cite{kryszewski2006positivity}.
-A loose upper bound, which may nonetheless be tighter than that in Eq.~\eqref{eq:bound} of the main text (derived in Appendix \ref{sec:bound}), might be found by the fact that physical qudit states $\rho$ satisfy $\obk{\rho|\rho}\le1$.
+  - \f1n \norm{\rho-\sigma_V^\star}_V^2,
+  &&
+  \norm{X}_V^2 \equiv \sum_\ell \bk{X_\ell|\N_{V\ell}|X_\ell},
+\end{align}
+where $X_\ell\equiv\sum_m\obk{T_{\ell m}|X}\ket{m}$ is a vector of the degree-$\ell$ components of $X$ in the polarization operator basis, and $\norm{X}_V$ is a noise-weighted norm of $X$.
+Maximizing $\E_V$ over all qudit states $\rho$ thus amounts to finding the closest physical qudit state $\rho$ to $\sigma_V^\star$, with distance measured by the metric $D_V\p{X,Y}\equiv\norm{X-Y}_V$.
+We leave this minimization problem to future work, and note that solving it will likely require making use of the positivity conditions derived in Ref.~\cite{kryszewski2006positivity}.
+A loose lower bound on $\norm{\rho-\sigma_V^\star}$ can be found by minimization under the constraint $\norm{\rho}\le1$ (or $\bk{\rho_\ell|\rho_\ell}\le1$), which may provide a tighter upper bound on $\E_V\p{\rho}$ than that in Eq.~\eqref{eq:bound} of the main text.
 

### 2020-11-23 sun_phases/spin_qudit_tomo.tex
refer to discussions about a tight reconstruction error bound in the main text
added discussion of the connection between polarization operators and spherical harmonics (via phase-space representations) added references for ultracold molecules

@@
 %%% symbols, notations, etc.
-\usepackage{physics,braket,bm,amssymb} % physics and math
+\usepackage{physics,braket,bm} % physics and math
 \renewcommand{\t}{\text} % text in math mode
@@
 
-% to place figures in the correct section
-\usepackage[section]{placeins}
-
 \usepackage[inline]{enumitem} % in-line lists and \setlist{} (below)
@@
 Quantum state tomography, or the task of reconstructing a quantum state by collecting and processing measurement data, is an essential primitive for quantum sensing, quantum simulation, and quantum information processing.
-The central importance of quantum state tomography has led to the development of a variety of techniques based on compressed sensing \cite{gross2010quantum}, matrix product states \cite{cramer2010efficient}, maximum-likelihood estimation \cite{smolin2012efficient}, Bayesian inference \cite{huszar2012adaptive}, and neural networks \cite{torlai2018neuralnetwork}, among others \cite{christandl2012reliable, qi2013quantum, opatrny1997leastsquares}.
+The central importance of quantum state tomography has led to the development of a variety of techniques based on compressed sensing \cite{gross2010quantum}, matrix product states \cite{cramer2010efficient}, maximum-likelihood estimation \cite{smolin2012efficient}, Bayesian inference \cite{huszar2012adaptive}, and neural networks \cite{torlai2018neuralnetwork}, among many others \cite{christandl2012reliable, qi2013quantum, opatrny1997leastsquares}
 These techniques are typically developed in a general, information-theoretic setting, and make minimal assumptions about the physical medium of a quantum state.
@@
 
-Due in part to advancements in experimental capabilities to address nuclear spins in ultracold atomic systems [\red{CITE}], as well as developments in molecular cooling techniques [\red{CITE}], a particular setting of growing interest is the spin qudit, or a multilevel quantum angular momentum degree of freedom.
+Due to advancements in experimental capabilities to address nuclear spins in ultracold atomic systems [\red{CITE}], as well as developments in the control of ultracold molecular systems \cite{zeppenfeld2012sisyphus, kozyryev2017sisyphus, puri2017synthesis, wu2017cryofuge, marco2019degenerate, liu2019molecular, anderegg2019optical, chou2020frequencycomb, lin2020quantum, covey2018approach, jamadagni2019quantum}, a particular setting of growing interest is the spin qudit, or a multilevel quantum angular momentum degree of freedom.
 Spin qudits can provide advantages over their qubit counterparts for quantum sensing \cite{evrard2019enhanced}, enable quantum simulations of SU($d$) magnetism \cite{banerjee2013atomic, cazalilla2014ultracold, zhang2014spectroscopic, scazza2014observation, goban2018emergence}, and offer unique capabilities for quantum error correction \cite{albert2020robust, gross2020encoding}.
@@
 
-The problem of qudit tomography in particular is not new, with an extensive literature on a variety of techniques \cite{manko1997spin, schmied2011tomographic, evrard2019enhanced, flammia2005minimal, thew2002qudit, salazar2012quantum, sosa-martinez2017quantum, ha2018minimal, stefano2019set, palici2020oam}.
-However, existing protocols either rely infinite-dimensional representations of a quantum spin \cite{manko1997spin, schmied2011tomographic, evrard2019enhanced}, or require the capability to perform essentially arbitrary operations on a qudit \cite{thew2002qudit, flammia2005minimal, salazar2012quantum, sosa-martinez2017quantum, ha2018minimal, stefano2019set, palici2020oam}, resulting in tomographic protocols that can be highly inefficient or unachievable in practice.
+The general problem of qudit tomography is not new, with an extensive literature on a variety of techniques \cite{manko1997spin, schmied2011tomographic, evrard2019enhanced, flammia2005minimal, thew2002qudit, salazar2012quantum, sosa-martinez2017quantum, ha2018minimal, stefano2019set, palici2020oam}.
+However, existing protocols either rely infinite-dimensional representations of a quantum spin \cite{manko1997spin, schmied2011tomographic, evrard2019enhanced}, or require the capability to perform essentially arbitrary operations on a qudit \cite{thew2002qudit, flammia2005minimal, salazar2012quantum, sosa-martinez2017quantum, ha2018minimal, stefano2019set, palici2020oam}, generally resulting in tomographic protocols that can be highly inefficient or unachievable in practice.
 The protocols based on infinite-dimensional representations of a quantum spin have the advantage of reconstructing its state from measurements of spin projection onto different spatial axes, which are generally accessible with any spin qudit.
-Nonetheless, these protocols obfuscate the actual requirements for performing full state tomography, provide no straightforward error bounds or guarantees of accuracy, and (with the exception of Ref.~\cite{schmied2011tomographic}) extract only a small faction of the information contained in measurement data.
+Nonetheless, these protocols obfuscate the minimal requirements for performing full state tomography, provide no straightforward error bounds or guarantees of accuracy, and (with the exception of Ref.~\cite{schmied2011tomographic}) extract only a small faction of the information contained in measurement data.
 
 In this work, we provide an efficient framework for performing spin qudit tomography using only measurements of spin projection onto different spatial axes.
-We first introduce, in Section \ref{sec:polarization_ops}, a set of qudit operators that have a simple physical interpretation, transform nicely under 3D rotations, and play a central role in our work.
-We then map the problem of spin qudit tomography onto the well-studied classical problem of signal recovery on the sphere in Section \ref{sec:signal_recovery}, thereby importing a host of existing literature and mathematical machinery that has been developed for this task \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
+We first introduce, in Section \ref{sec:polarization_ops}, a set of qudit operators that are a quantum analogue of the spherical harmonics, and which play a central role in our work.
+We then map the quantum problem of spin qudit tomography onto the well-studied classical problem of signal recovery on the sphere in Section \ref{sec:signal_recovery}, thereby importing a host of existing literature and mathematical machinery that has been developed for this task \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
 In particular, this mapping allows us to bound the minimum number of measurement axes necessary to perform full tomography on a $d$-level spin qudit, $r_d^{\t{min}}$, by $2d-1\le r_d^{\t{min}}\le d^2$.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Polarization operators and 3D rotations}
+\section{Polarization operators}
 \label{sec:polarization_ops}
 
+We begin by introducing a set of qudit operators that are analogous to the spherical harmonics $Y_{\ell m}$, and which play a central role in our work.
 Consider a $d$-state spin qudit with total spin $s\equiv\frac{d-1}{2}$.
-The defining property of a spin qudit is the fact that it describes an angular momentum degree of freedom, which has specific implications for how a spin qudit should transform under the group SO(3) of rotations in 3D space.
-Due to the central importance of these transformation rules for a spin qudit, we first identify a basis of operators that transforms nicely under 3D rotations.
-One such basis is provided by the {\it polarization operators} \cite{kryszewski2006positivity, bertlmann2008bloch}, defined by
+The defining property of a spin qudit is that it describes an angular momentum degree of freedom, which has specific implications for how a spin qudit should transform under the group SO(3) of rotations in 3D space.
+Due to the central importance of these transformation rules for a spin qudit, we seek a basis of operators that transform nicely under 3D rotations \footnote{Technically speaking, we seek a basis of operators that transform as an irreducible representation of SO(3).}.
+One such basis is that of the {\it polarization operators} \cite{kryszewski2006positivity, bertlmann2008bloch}, defined by
 \begin{align}
@@
 \end{align}
-where $\bk{s\mu;\ell m|s\nu}$ is a Clebsh-Gordan coefficient that enforces $\ell\in\set{0,1,\cdots,d-1}$ and $m\in\set{-\ell,-\ell+1,\cdots,\ell}$, such that there are $d^2$ operators in total.
+where $\bk{s\mu;\ell m|s\nu}$ is a Clebsh-Gordan coefficient that enforces $\ell\in\set{0,1,\cdots,d-1}$ and $m\in\set{-\ell,-\ell+1,\cdots,\ell}$, such that there are $d^2$ polarization operators in total.
 For brevity, we will generally treat the value of $d$ (or $s$) as constant but arbitrary throughout this work, and suppress any explicit dependence of quantities or operators such as $T_{\ell m}$ on $d$.
@@
 where $\bk{X}_\rho\equiv\tr\p{\rho X}=\obk{\rho|X}$, and $\rho^\dag=\rho$ implies that $\rho_{\ell m}^*=\p{-1}^m\rho_{\ell,-m}$.
-The polarization operators can be physically interpreted as the polarization $\ket\psi\to T_{\ell m}\ket\psi$ induced on the state $\ket\psi$ of a spin-$s$ system by the absorption of a spin-$\ell$ particle with spin projection $m$ onto a fixed quantization axis.
+The polarization operators can be physically interpreted by the transition $\ket\psi\to T_{\ell m}\ket\psi$ induced on the state $\ket\psi$ of a spin-$s$ system by the absorption of a particle with total spin $\ell$ and spin projection $m$ onto a fixed quantization axis.
 In analogy with the spherical harmonics $Y_{\ell m}$, we will refer to the index $\ell$ as the {\it degree} and $m$ as the {\it order} of $T_{\ell m}$.
@@
 \end{align}
-and expand rotations of the polarization operators as
+and expand polarization operators in a rotated frame as
 \begin{align}
   T_{\v\omega\ell m} \equiv
-  R\p{\v\omega} T_{\ell m} R\p{\v\omega}^\dag
+  R\p{\v\omega}^\dag T_{\ell m} R\p{\v\omega}
   = \sum_{n=-\ell}^\ell D_{nm}^\ell\p{\v\omega} T_{\ell n},
@@
 are (Wigner) rotation matrix elements.
-As a final comment, we note that the polarization operators are closely related to the ``Weyl-Wigner-Moyal'' formalism for spin qudits \cite{li2013weylwignermoyal}, which provides spherical tensor operators with a phase-space representation that is essentially equal to the spherical harmonics $Y_{\ell m}$ (up to proportionality factors determined by $\ell$ and $m$).
-The precise connection between this formalism and polarization operators would be interesting to clarify in future work.
-
-Throughout this work, we will primarily consider rotations of the sphere that take the north pole to a fixed point $\v v=\p{\alpha,\beta}$ at azimuthal angle $\alpha$ and polar angle $\beta$.
+Throughout this work, we will primarily consider rotations of the sphere that take the north pole to a point $\v v=\p{\alpha,\beta}$ at azimuthal angle $\alpha$ and polar angle $\beta$.
 For ease of notation, we will therefore implicitly associate Euler angle doublets $\v v=\p{\alpha,\beta}$ with the triplet $\p{\alpha,\beta,0}$, such that $R\p{\v v} \equiv R\p{\alpha,\beta,0}$, $D^\ell_{mn}\p{\v v} \equiv D^\ell_{mn}\p{\alpha,\beta,0}$, and $T_{\v v\ell m} \equiv T_{\p{\alpha,\beta,0},\ell m}$.
-We will also primarily consider polarization operators with order $m=0$.
-Symbols with a suppressed order therefore implicitly set this order to zero, which is to say that $T_\ell\equiv T_{\ell,0}$, $T_{\v v\ell}\equiv T_{\v v\ell,0}$, and $D^\ell_n\equiv D^\ell_{n,0}$.
+We will also frequently consider polarization operators with order $m=0$.
+Symbols with a suppressed order therefore implicitly set this order to zero, which is to say that $T_\ell\equiv T_{\ell,0}$, $T_{\v v\ell}\equiv T_{\v v\ell,0}$, and $D^\ell_m\equiv D^\ell_{m,0}$.
+
+As a final point, we mention that the polarization operators $T_{\ell m}$ share a connection to the spherical harmonics $Y_{\ell m}$ that goes beyond the rules for their transformation under 3D rotations.
+In fact, the phase-space (Wigner function) representation of $T_{\ell m}$ is essentially equal to $Y_{\ell m}$ (up to complex conjugation and a normalization factor).
+The phase-space representation of an operator $X$ assigns, to each point $\v v$ on the sphere, the complex number
+\begin{align}
+  \tilde X\p{\v v} \equiv \bk{s_{\v v}|X|s_{\v v}},
+\end{align}
+where $\ket{s_{\v v}}\equiv R\p{\v v}\ket{s}$ is the state of a spin qudit polarized along $\v v$.
+This representation is faithful in the sense that $X$ is uniquely determined by the phase-space values $\tilde X\p{\v v}$ at all points $\v v$ on the sphere.
+The transformation rules for polarization operators in Eq.~\eqref{eq:trans_rot}, together with the fact that $\bk{s|T_{\ell m}|s}=0$ unless $m=0$, suffice to show that their phase-space distributions are
+\begin{align}
+  \tilde T_{\ell m}\p{\v v} = c_\ell Y_{\ell m}\p{\v v}^*,
+\end{align}
+where the proportionality factor $c_\ell$ simply enforces that $\obk{T_{\ell m}|T_{\ell m}}=1$ (see Appendix \ref{sec:rotations}).
+In this way, the polarization operators $T_{\ell m}$ are a quantum analogue of spherical harmonics $Y_{\ell m}$, which play an important role in phase-space formalisms for spin qudits \cite{li2013weylwignermoyal}.
 
@@
 Our goal is to reconstruct an arbitrary state $\rho$ of a spin qudit from measurements of spin projection onto different quantization axes.
-We are thus nominally restricted to measuring projectors $\Pi_{\v v\mu} \equiv \op{\mu_{\v v}}$, where $\ket{\mu_{\v v}}\equiv R\p{\v v}\ket{\mu}$ is a state with spin projection $\mu$ onto a measurement axis $\v v$ determined by a point on the sphere.
-For any fixed axis $\v v$, the sets $\set{\Pi_{\v v\mu}}$ and $\set{T_{\v v\ell}}$ are both orthonormal bases for the space of operators that are diagonal in the basis $\set{\ket{\mu_{\v v}}}$.
-By a simple change of operator basis, measuring the projectors $\set{\Pi_{\v v\mu}}$ is therefore equivalent to measuring the polarization operators $\set{T_{\v v\ell}}$, and provides data on the expectation values $\bk{T_{\v v\ell}}_\rho$.
-
-Reconstructing an arbitrary density operator $\rho$ essentially requires us to find a set of axes $V=\set{\v v}$ and associated coefficients $C_{\ell mk}\p{\v v}$ that would allow us to recover any coefficient $\rho_{\ell m}$ in the expansion in Eq.~\eqref{eq:trans_state} through
+We are thus nominally restricted to measuring projectors $\Pi_{\v v\mu} \equiv \op{\mu_{\v v}}$, where $\ket{\mu_{\v v}}\equiv R\p{\v v}\ket{\mu}$ is a state with spin projection $\mu$ onto the measurement axis $\v v$.
+For any fixed axis $\v v$, the sets $\set{\Pi_{\v v\mu}}$ and $\set{T_{\v v\ell}}$ are both complete bases for the space of operators that are diagonal in the basis $\set{\ket{\mu_{\v v}}}$.
+Measuring the projectors $\set{\Pi_{\v v\mu}}$ is therefore equivalent to measuring the polarization operators $\set{T_{\v v\ell}}$, and provides data on the expectation values $\bk{T_{\v v\ell}}_\rho$.
+
+In order to reconstruct an arbitrary density operator $\rho$ from the expectation values $\bk{T_{\v v\ell}}_\rho$, we essentially need to to find a set of coefficients $C_{\ell mk}\p{\v v}$ that would allow us to recover any matrix element $\rho_{\ell m}$ of $\rho$ through
 \begin{align}
@@
 In fact, the problem of finding suitable axes $V$ and coefficients $C_{\ell m}\p{\v v}$ to satisfy Eq.~\eqref{eq:tomo_recovery}, solving which would enable the recovery of arbitrary spin qudit states, can be mapped onto the well-studied problem of signal recovery on the sphere \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
-The signal recovery problem can be stated as follows: given a square-integrable function $f$ on the sphere, with the harmonic expansion
+The signal recovery problem can be stated as follows: given a square-integrable function $f$ on the sphere, with the spherical harmonic expansion
 \begin{align}
@@
 \end{align}
-where $f_{\ell m}$ are expansion coefficients and $Y_{\ell m}$ is the spherical harmonic of degree $\ell$ and order $m$, find a set of points $V=\set{\v v}$ and associated coefficients $\tilde C_{\ell m}\p{\v v}$ with which we can reconstruct $f$, or equivalently its coefficients $f_{\ell m}$, from knowledge of the function's value $f\p{\v v}$ at all points $\v v\in V$; that is
+where $f_{\ell m}$ are complex coefficients, find a set of points $V=\set{\v v}$ and associated coefficients $\tilde C_{\ell m}\p{\v v}$ with which we can reconstruct $f$, or equivalently its coefficients $f_{\ell m}$, from knowledge of the function's value $f\p{\v v}$ at all points $\v v\in V$; that is
 \begin{align}
@@
   For a $d$-level spin, these quasi-probability distributions are band-limited at degree $d$.}.
-Moreover, for any fixed degree $\ell$, finding solutions to the reduced recovery problem in Eq.~\eqref{eq:tomo_recovery} is equivalent to the recovery of a degree-$\ell$ function $f = \sum_m f_m Y_{\ell m}$, which is provably possible with $\abs{V}=2\ell+1$ samples \cite{freeden2008spherical}.
+Moreover, for any fixed degree $\ell$, finding solutions to the reduced recovery problem in Eq.~\eqref{eq:tomo_recovery} is equivalent to the recovery of a degree-$\ell$ function $f_\ell = \sum_m f_{\ell m} Y_{\ell m}$, which is provably possible with $\abs{V}=2\ell+1$ samples \cite{freeden2008spherical}.
 In the case of spin qudit tomography, the degree $\ell$ takes a maximal value of $\ell_{\t{max}}\equiv d-1$, so state recovery requires at least as many measurement axes as there are polarization operators with degree $\ell_{\t{max}}$, namely $2\ell_{\t{max}}+1=2d-1$.
@@
 one might expect that $2d-1$ suitably chosen measurement axes would always suffice to perform full tomography of a $d$-level spin qudit.
-In the language of classical signal recovery, we would expect there to exist a set of $2\ell_{\t{max}}+1$ points on the sphere that enable the recovery of {\it any} function with fixed degree $\ell\le\ell_{\t{max}}$ (given prior knowledge of $\ell$) \footnote{We could further conjecture that if {\it any} set of points $V\subset\Omega_2$ enables the recovery of degree-$\ell_{\t{max}}$ functions on the sphere, then $V$ enables the recovery degree-$\ell$ functions with $\ell\le\ell_{\t{max}}$.
+In the language of classical signal recovery, we would expect there to exist a set of $2\ell_{\t{max}}+1$ points on the sphere that enable the recovery of {\it any} function with fixed degree $\ell\le\ell_{\t{max}}$, given prior knowledge of $\ell$ \footnote{We could further conjecture that if {\it any} set of points $V\subset\Omega_2$ enables the recovery of degree-$\ell_{\t{max}}$ functions on the sphere, then $V$ enables the recovery degree-$\ell$ functions with $\ell\le\ell_{\t{max}}$.
   However, this condition is stronger than necessary to guarantee full recovery of a $d$-level spin qudit state with $2d-1$ measurement axes, which only requires the existence of {\it some} such set of points.}.
@@
 By a simple change of basis, measuring these projectors is equivalent to measuring the polarization operators $\set{T_{\v v\ell}}$.
-By flattening each $d\times d$ matrix $T_{\v v\ell}$ into the $d^2$-component column vector $\oket{T_{\v v\ell}}$, we can construct the {\it measurement matrix}
+Flattening each $d\times d$ matrix $T_{\v v\ell}$ into the $d^2$-component column vector $\oket{T_{\v v\ell}}$, we construct the {\it measurement matrix}
 \begin{align}
@@
 There is therefore still room for improvement on the bound in Eq.~\eqref{eq:bound} by maximizing $\E_V$ over the set of all qudit states $\rho$.
+We discuss this maximization problem in Appendix \ref{sec:exact}.
 We also note that the (squared) reconstruction error bound in Eq.~\eqref{eq:bound} obeys the ``standard quantum limit'' of $\sim1/n$ scaling in the number of measurements.
@@
 The actual error in the reconstruction $\tilde\rho_V$ of a particular state $\rho$ may be considerably smaller, and depends on $\rho$ itself.
-The exact mean squared distance between $\tilde\rho_V$ and $\rho$, derived in Appendix \ref{sec:exact_error}, can be written in the form
+The exact mean squared distance between $\tilde\rho_V$ and $\rho$, derived in Appendix \ref{sec:exact}, can be written in the form
 \begin{align}
@@
 
-\begin{figure}
+The reconstruction error bound in Eq.~\eqref{eq:bound} nominally provides a straightforward answer: at a fixed total number of measurements, $N=n\abs{V}$, the number of measurement axes should be chosen to minimize the (squared) reconstruction error $\E_V\p{\rho}^2<\epsilon_V^2/n\propto\epsilon_V^2\abs{V}$.
+We therefore consider the measurement-adjusted error scale
+\begin{align}
+  \beta(p)
+  \equiv \min\set{\epsilon_V:\abs{V}=2d-1+p} \times \sqrt{2d-1+p}
+\end{align}
+as a function of $p$, the number of measurement axes exceeding $2d-1$.
+Though we cannot minimize over all suitable choices of measurement axes $V$ to compute $\beta(p)$, we can compute an empirical upper bound $\tilde\beta(p)\ge\beta(p)$ by minimizing over a large number of randomly chosen $V$.
+Figure \ref{fig:axes} shows the results of such empirical minimization, where we find that $\tilde\beta(p)$ drops substantially before plateauing at $p\approx d$, after which there are only marginal benefits to using more measurement axes.
+In the interest of reducing experimental complexity as well as the runtime of our tomography protocol, which grow linearly in $\abs{V}$, we therefore conclude that spin qudit tomography should be performed with $\abs{V}\approx3d$ measurement axes.
+
+\begin{figure}[h]
   \centering
@@
 
-The reconstruction error bound in Eq.~\eqref{eq:bound} nominally provides a straightforward answer: at a fixed total number of measurements, $N=n\abs{V}$, the number of measurement axes should be chosen to minimize the (squared) reconstruction error $\E_V\p{\rho}^2<\epsilon_V^2/n\propto\epsilon_V^2\abs{V}$.
-We therefore consider the measurement-adjusted error scale
-\begin{align}
-  \beta(p)
-  \equiv \min\set{\epsilon_V:\abs{V}=2d-1+p} \times \sqrt{2d-1+p}
-\end{align}
-as a function of $p$, the number of measurement axes exceeding $2d-1$.
-Though we cannot minimize over all suitable choices of measurement axes $V$ to compute $\beta(p)$, we can compute an empirical upper bound $\tilde\beta(p)\ge\beta(p)$ by minimizing over a large number of randomly chosen $V$.
-Figure \ref{fig:axes} shows the results of such empirical minimization, where we find that $\tilde\beta(p)$ drops substantially before plateauing at $p\approx d$, after which there are only marginal benefits to using more measurement axes.
-In the interest of reducing experimental complexity as well as the runtime of our tomography protocol, which grow linearly in $\abs{V}$, we therefore conclude that spin qudit tomography should be performed with $\abs{V}\approx3d$ measurement axes.
-
 \bibliography{\jobname.bib}
@@
 Moreover, by comparing Eqs.~\eqref{eq:spin_ops} and \eqref{eq:spin_trans} we see that the polarization operators $T_{\ell m}$ transform identically to spin-$\ell$ particles under the adjoint action of the spin operators $S_\z$ and $S_\pm$.
-For any triplet of Euler angles $\v\omega=\p{\alpha,\beta,\gamma}$, we can therefore define the fundamental rotation operator
+For any triplet of Euler angles $\v\omega=\p{\alpha,\beta,\gamma}$, we can therefore define the rotation operator
 \begin{align}
@@
 \end{align}
-and expand
+and expand polarization operators in a rotated frame as
 \begin{align}
   T_{\v\omega\ell m}
-  \equiv R\p{\v\omega} T_{\ell m} R\p{\v\omega}^\dag
+  \equiv R\p{\v\omega}^\dag T_{\ell m} R\p{\v\omega}
   = \sum_{n=-\ell}^\ell D_{nm}^\ell\p{\v\omega} T_{\ell n},
+  \label{eq:trans_rot_apndx}
 \end{align}
@@
   D_{mn}^\ell\p{\v\omega} \equiv \bk{\ell m|R\p{\v\omega}|\ell n}
-  = \obk{T_{\ell m}|R\p{\v\omega}\otimes R\p{\v\omega}^*|T_{\ell n}}
+  = \obk{T_{\ell m} | R\p{\v\omega}^\dag
+    \otimes R\p{\v\omega}^* |T_{\ell n}}
 \end{align}
 are matrix elements of the rotation operator $R\p{\v\omega}$ for spin-$\ell$ particles.
+
+Implicitly associating the Euler angle doublet $\v v\equiv\p{\alpha,\beta}$ with the triplet $\p{\alpha,\beta,0}$, the transformation rules in Eq.~\eqref{eq:trans_rot_apndx} imply that we can expand the phase-space distribution of $T_{\ell m}$ as
+\begin{align*}
+  \tilde T_{\ell m}\p{\v v}
+  \equiv \bk{s_{\v v} | T_{\ell m} | s_{\v v}}
+  = \bk{s | R\p{\v v}^\dag T_{\ell m} R\p{\v v} | s}
+  = D^\ell_{0,m}\p{\v v} \bk{s | T_{\ell,0} | s},
+\end{align*}
+where
+\begin{align*}
+  \bk{s | T_{\ell,0} | s}
+  = \sqrt{\f{2\ell+1}{2s+1}} \, \bk{ss;\ell,0|ss}
+  = \sqrt{\f{2\ell+1}{2s+\ell+1}
+    \p{\f{\p{2s}!}{\p{2s+\ell}!}}
+    \p{\f{\p{2s}!}{\p{2s-\ell}!}}},
+\end{align*}
+and the properties of the rotation matrix elements $D^\ell_{mn}$ imply that
+\begin{align*}
+  D^\ell_{0,m}\p{\v v} = \p{-1}^m D^\ell_{-m,0}\p{\v v}^*
+  = \sqrt{\f{4\pi}{2\ell+1}} \p{-1}^m Y_{\ell,-m}\p{\v v}
+  = \sqrt{\f{4\pi}{2\ell+1}} \, Y_{\ell m}\p{\v v}^*,
+\end{align*}
+so
+\begin{align*}
+  \tilde T_{\ell m}\p{\v v}
+  = \sqrt{\f{4\pi}{2s+\ell+1}
+    \p{\f{\p{2s}!}{\p{2s+\ell}!}} \p{\f{\p{2s}!}{\p{2s-\ell}!}}} \,
+  Y_{\ell m}\p{\v v}^*
+  \propto Y_{\ell m}\p{\v v}^*.
+\end{align*}
+In this way, the polarization operators are a quantum analogue of spherical harmonics.
 
@@
 \section{Exact reconstruction error}
-\label{sec:exact_error}
+\label{sec:exact}
 
@@
 \subsection{Comments on a tight reconstruction error bound}
+\label{sec:tight}
 

### 2020-11-23 sun_phases/spin_qudit_tomo.tex
minor formatting changes, placing footnotes in the main text

@@
-\documentclass[notitlepage,twocolumn]{revtex4-2}
+\documentclass[notitlepage,nofootinbib,twocolumn]{revtex4-2}
 
@@
 
-Due to advancements in experimental capabilities to address nuclear spins in ultracold atomic systems [\red{CITE}], as well as developments in the control of ultracold molecular systems \cite{zeppenfeld2012sisyphus, kozyryev2017sisyphus, puri2017synthesis, wu2017cryofuge, marco2019degenerate, liu2019molecular, anderegg2019optical, chou2020frequencycomb, lin2020quantum, covey2018approach, jamadagni2019quantum}, a particular setting of growing interest is the spin qudit, or a multilevel quantum angular momentum degree of freedom.
-Spin qudits can provide advantages over their qubit counterparts for quantum sensing \cite{evrard2019enhanced}, enable quantum simulations of SU($d$) magnetism \cite{banerjee2013atomic, cazalilla2014ultracold, zhang2014spectroscopic, scazza2014observation, goban2018emergence}, and offer unique capabilities for quantum error correction \cite{albert2020robust, gross2020encoding}.
+Due to advancements in experimental capabilities to address nuclear spins in ultracold atomic systems \cite{cazalilla2014ultracold} [\red{more references?}], as well as developments in the control of ultracold molecular systems \cite{zeppenfeld2012sisyphus, kozyryev2017sisyphus, puri2017synthesis, wu2017cryofuge, marco2019degenerate, liu2019molecular, anderegg2019optical, chou2020frequencycomb, lin2020quantum, covey2018approach, jamadagni2019quantum}, a particular setting of growing interest is the spin qudit, or a multilevel quantum angular momentum degree of freedom.
+Spin qudits can provide advantages over their qubit counterparts for quantum sensing \cite{evrard2019enhanced}, enable quantum simulations of SU($d$) magnetism \cite{cazalilla2014ultracold, banerjee2013atomic, zhang2014spectroscopic, scazza2014observation, goban2018emergence}, and offer unique capabilities for quantum error correction \cite{albert2020robust, gross2020encoding}.
 In all cases, quantum state tomography is necessary to take full advantage of a spin qudit.
@@
 The defining property of a spin qudit is that it describes an angular momentum degree of freedom, which has specific implications for how a spin qudit should transform under the group SO(3) of rotations in 3D space.
-Due to the central importance of these transformation rules for a spin qudit, we seek a basis of operators that transform nicely under 3D rotations \footnote{Technically speaking, we seek a basis of operators that transform as an irreducible representation of SO(3).}.
+Due to the central importance of these transformation rules for a spin qudit, we seek a basis of operators that transform nicely under 3D rotations.\footnote{Technically speaking, we seek a basis of operators that transform as an irreducible representation of SO(3).}
 One such basis is that of the {\it polarization operators} \cite{kryszewski2006positivity, bertlmann2008bloch}, defined by
@@
 If the function $f$ is {\it band-limited} at degree $L$, which is to say that $f_{\ell m}=0$ for all $\ell\ge L$, then the full recovery problem in Eq.~\eqref{eq:full_recovery} is provably solvable with a suitable choice of $\abs{V}=L^2$ points on the sphere \cite{freeden2008spherical, freeden2018spherical}.
-The existence of these solutions to the full recovery problem in turn implies the existence of $d^2$ measurement axes that suffice to reconstruct arbitrary states of $d$-level spin qudit, whose possible states are essentially ``band-limited'' at degree $d$ \footnote{Spin qudit states can be faithfully represented by quasi-probability distributions on a sphere.
-  For a $d$-level spin, these quasi-probability distributions are band-limited at degree $d$.}.
+The existence of these solutions to the full recovery problem in turn implies the existence of $d^2$ measurement axes that suffice to reconstruct arbitrary states of $d$-level spin qudit, whose possible states are essentially ``band-limited'' at degree $d$.\footnote{Spin qudit states can be faithfully represented by quasi-probability distributions on a sphere.
+  For a $d$-level spin, these quasi-probability distributions are band-limited at degree $d$.}
 Moreover, for any fixed degree $\ell$, finding solutions to the reduced recovery problem in Eq.~\eqref{eq:tomo_recovery} is equivalent to the recovery of a degree-$\ell$ function $f_\ell = \sum_m f_{\ell m} Y_{\ell m}$, which is provably possible with $\abs{V}=2\ell+1$ samples \cite{freeden2008spherical}.
@@
 one might expect that $2d-1$ suitably chosen measurement axes would always suffice to perform full tomography of a $d$-level spin qudit.
-In the language of classical signal recovery, we would expect there to exist a set of $2\ell_{\t{max}}+1$ points on the sphere that enable the recovery of {\it any} function with fixed degree $\ell\le\ell_{\t{max}}$, given prior knowledge of $\ell$ \footnote{We could further conjecture that if {\it any} set of points $V\subset\Omega_2$ enables the recovery of degree-$\ell_{\t{max}}$ functions on the sphere, then $V$ enables the recovery degree-$\ell$ functions with $\ell\le\ell_{\t{max}}$.
-  However, this condition is stronger than necessary to guarantee full recovery of a $d$-level spin qudit state with $2d-1$ measurement axes, which only requires the existence of {\it some} such set of points.}.
+In the language of classical signal recovery, we would expect there to exist a set of $2\ell_{\t{max}}+1$ points on the sphere that enable the recovery of {\it any} function with fixed degree $\ell\le\ell_{\t{max}}$, given prior knowledge of $\ell$.\footnote{We could further conjecture that if {\it any} set of points $V\subset\Omega_2$ enables the recovery of degree-$\ell_{\t{max}}$ functions on the sphere, then $V$ enables the recovery degree-$\ell$ functions with $\ell\le\ell_{\t{max}}$.
+  However, this condition is stronger than necessary to guarantee full recovery of a $d$-level spin qudit state with $2d-1$ measurement axes, which only requires the existence of {\it some} such set of points.}
 Unfortunately, we have no general proof of this conjecture, which requires a deeper analysis of fixed-degree signal recovery on the sphere.
@@
 
-The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and optimize (using any standard minimization algorithm) the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the quantum error scale $\epsilon_V$.
+The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere,\footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.} and optimize (using any standard minimization algorithm) the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the quantum error scale $\epsilon_V$.
 If $\abs{V}$ is too large for such optimization to be practical, one can instead generate many random sets of measurement axes, and then choose the set with the smallest quantum error scale $\epsilon_V$.
@@
 
-\begin{figure}[h]
+\begin{figure}
   \centering

### 2020-11-23 sun_phases/spin_qudit_tomo.tex
fix minor error

@@
   = \obk{T_{\ell m} | R\p{\v\omega}^\dag
-    \otimes R\p{\v\omega}^* |T_{\ell n}}
-\end{align}
-are matrix elements of the rotation operator $R\p{\v\omega}$ for spin-$\ell$ particles.
-
-Implicitly associating the Euler angle doublet $\v v\equiv\p{\alpha,\beta}$ with the triplet $\p{\alpha,\beta,0}$, the transformation rules in Eq.~\eqref{eq:trans_rot_apndx} imply that we can expand the phase-space distribution of $T_{\ell m}$ as
+    \otimes R\p{\v\omega}^{\t{T}} |T_{\ell n}}
+\end{align}
+are matrix elements of the rotation operator $R\p{\v\omega}$ for spin-$\ell$ particles, and $R\p{\v\omega}^{\t{T}}$ is the transpose of $R\p{\v\omega}$.
+
+Implicitly associating the Euler angle doublet $\v v = \p{\alpha,\beta}$ with the triplet $\p{\alpha,\beta,0}$, the transformation rules in Eq.~\eqref{eq:trans_rot_apndx} imply that we can expand the phase-space distribution of $T_{\ell m}$ as
 \begin{align*}

### 2020-11-23 sun_phases/spin_qudit_tomo.tex
made the tomography protocol more explicit

--- 2020-11-23 sun_phases/spin_qudit_tomo.tex
@@
 In this case $M_V$ must be full rank, with $d^2$ nonzero singular values.
-Indexing these singular values $M^V_k$ and the corresponding (normalized) right singular vectors $\v x^V_k\equiv(x^V_{k,1},x^V_{k,2},\cdots,x^V_{k,\abs{V}})$ by an integer $k\in\set{1,2,\cdots,d^2}$, we can construct the orthonormal qudit operators
+Indexing these singular values $M^V_k$ and the corresponding (normalized) right singular vectors $\v x^V_k\equiv \sum_j x^V_{kj} \ket{j}$ by an integer $k\in\set{1,2,\cdots,d^2}$, we can construct the orthonormal qudit operators
 \begin{align}
@@
   &&
-  \v q^V_k \equiv \f{\v x^V_k}{M^V_k},
+  q^V_{kj} \equiv \f{x^V_{kj}}{M^V_k},
 \end{align}
@@
   = \sum_{m,\v v} D^\ell_m\p{\v v} \op{m}{\v v}.
+  \label{eq:block}
 \end{align}
@@
 If $\abs{V}$ is too large for such optimization to be practical, one can instead generate many random sets of measurement axes, and then choose the set with the smallest quantum error scale $\epsilon_V$.
-After choosing a set of measurement axes $V$, (experimentally) make $n$ measurements of spin projection along every axis $\v v\in V$, and set $\tilde\Pi_{\v v\mu}\approx\bk{\Pi_{\v v\mu}}_\rho$ to the fraction of times in which the measurement outcome was $\mu$.
-The estimates $\tilde\Pi_{\v v\mu}$ of $\bk{\Pi_{\v v\mu}}_\rho$ determine estimates $\tilde T_{\v v\ell} \equiv \sum_\mu \bk{\mu|T_\ell|\mu}\tilde\Pi_{\v v\mu}$ of $\bk{T_{\v v\ell}}_\rho$, which can in turn be used to construct the estimate $\tilde\rho_V$ of $\rho$.
+After choosing a set of measurement axes $V$, make $n$ measurements of spin projection along every axis $\v v\in V$, and set $\tilde\Pi_{\v v\mu}\approx\bk{\Pi_{\v v\mu}}_\rho$ to the fraction of times in which the measurement outcome was $\mu$.
+The estimates $\tilde\Pi_{\v v\mu}$ of $\bk{\Pi_{\v v\mu}}_\rho$ determine estimates
+\begin{align}
+  \tilde T_{\v v\ell}
+  \equiv \sum_\mu \bk{\mu|T_\ell|\mu} \tilde\Pi_{\v v\mu}
+\end{align}
+of $\bk{T_{\v v\ell}}_\rho$, which can in turn be used to construct the estimate $\tilde\rho_V$ of $\rho$.
+Specifically, the $2\ell+1$ (normalized) right singular vectors $\v x^V_{\ell k} = \sum_{\v v} x^V_{\ell k \v v} \ket{\v v}$ and corresponding singular values $M^V_{\ell k}$ of the matrices $M_{V\ell}$ defined in Eq.~\eqref{eq:block} determine the components
+\begin{align}
+  \tilde\rho^V_{\ell k}
+  = \f1{M^V_{\ell k}} \sum_{\v v}
+  x^V_{\ell k\v v} \, \tilde T_{\v v\ell},
+\end{align}
+of
+\begin{align}
+  \tilde\rho_V = \sum_{\ell,k} \tilde\rho^V_{\ell k} Q_{\ell k}
+  \approx \rho,
+  &&
+  Q_{\ell k} = \f1{M^V_{\ell k}} \sum_{\v v}
+  x^V_{\ell k\v v} T_{\v v\ell}.
+\end{align}
 The total error in $\tilde\rho_V$, measured by its root-mean-square distance from $\rho$, is approximately $\E_V\p{\tilde\rho_V}\approx\E_V\p{\rho}$.
@@
 
-\begin{figure}[h]
+\begin{figure}[hb]
   \centering
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \section{Conclusions}
+
+% todo: write conclusions
+
 \bibliography{\jobname.bib}
@@
   = \obk{T_{\ell m} | R\p{\v\omega}^\dag
-    \otimes R\p{\v\omega}^* |T_{\ell n}}
-\end{align}
-are matrix elements of the rotation operator $R\p{\v\omega}$ for spin-$\ell$ particles.
-
-Implicitly associating the Euler angle doublet $\v v\equiv\p{\alpha,\beta}$ with the triplet $\p{\alpha,\beta,0}$, the transformation rules in Eq.~\eqref{eq:trans_rot_apndx} imply that we can expand the phase-space distribution of $T_{\ell m}$ as
-\begin{align*}
+    \otimes R\p{\v\omega}^{\t{T}} |T_{\ell n}}
+\end{align}
+are matrix elements of the rotation operator $R\p{\v\omega}$ for spin-$\ell$ particles, and $R\p{\v\omega}^{\t{T}}$ is the transpose of $R\p{\v\omega}$.
+
+Implicitly associating the Euler angle doublet $\v v = \p{\alpha,\beta}$ with the triplet $\p{\alpha,\beta,0}$, the transformation rules in Eq.~\eqref{eq:trans_rot_apndx} imply that we can expand the phase-space distribution of $T_{\ell m}$ as
+\begin{align}
   \tilde T_{\ell m}\p{\v v}
@@
   = D^\ell_{0,m}\p{\v v} \bk{s | T_{\ell,0} | s},
-\end{align*}
+\end{align}
 where
-\begin{align*}
+\begin{align}
   \bk{s | T_{\ell,0} | s}
@@
     \p{\f{\p{2s}!}{\p{2s-\ell}!}}},
-\end{align*}
+\end{align}
 and the properties of the rotation matrix elements $D^\ell_{mn}$ imply that
-\begin{align*}
+\begin{align}
   D^\ell_{0,m}\p{\v v} = \p{-1}^m D^\ell_{-m,0}\p{\v v}^*
@@
   = \sqrt{\f{4\pi}{2\ell+1}} \, Y_{\ell m}\p{\v v}^*,
-\end{align*}
+\end{align}
 so
-\begin{align*}
+\begin{align}
   \tilde T_{\ell m}\p{\v v}
@@
   \propto Y_{\ell m}\p{\v v}^*.
-\end{align*}
+\end{align}
 In this way, the polarization operators are a quantum analogue of spherical harmonics.
@@
 
-To this end, we define the probability $p^i_\mu\equiv\bk{\Pi_{i\mu}}_\rho$, collect these probabilities into the classical probability distribution $p^i = \p{p^i_{-s}, p^i_{-s+1}, \cdots, p^i_s}$, and define the vector $t_\ell \equiv \p{t_{\ell,-s},t_{\ell,-s+1},\cdots,t_{\ell s}}$ of all $t_{\ell\mu}\equiv\bk{\mu|T_\ell|\mu}$.
+To this end, we define the probability $p^i_\mu\equiv\bk{\Pi_{i\mu}}_\rho$, collect these probabilities into the classical probability distribution $p^i = \sum_\mu p^i_\mu \ket{\mu}$, and define the vector $t_\ell \equiv \sum_\mu t_{\ell\mu} \ket{\mu}$ of all $t_{\ell\mu}\equiv\bk{\mu|T_\ell|\mu}$.
 We then observe that
@@
 where $D^\ell_{mn}\p{\v v}\equiv\bk{\ell m|R\p{\v v}|\ell n}$ is a (Wigner) rotation matrix element for a spin-$\ell$ particle.
-The block-diagonal structure of $M_V$ allows us to index its singular values $M^V_{\ell m}$ and corresponding (normalized) right singular vectors $\v x^V_{\ell m} = (x^V_{\ell m,1}, x^V_{\ell m,2}, \cdots )$ by the indices $\p{\ell,m}$, where the integer $\abs{m}\le\ell$.
+The block-diagonal structure of $M_V$ allows us to index its singular values $M^V_{\ell m}$ and corresponding (normalized) right singular vectors $\v x^V_{\ell m} = \sum_i x^V_{\ell m i} \ket{\v v_i}$ by the indices $\p{\ell,m}$, where the integer $\abs{m}\le\ell$.
 These singular vectors and values define the orthonormal operators
@@
   &&
-  \v q^V_{\ell m} \equiv \f{\v x^V_{\ell m}}{M^V_{\ell m}},
-\end{align}
-where $i\in\set{1,2,\cdots,\abs{V}}$ indexes an axis $\v v_i\in V$.
+  q^V_{\ell m i} \equiv \f{x^V_{\ell m i}}{M^V_{\ell m}},
+\end{align}
+where $i\in\set{1,2,\cdots,\abs{V}}$ indexes an axis $\v v_i\in V$, with $T_{i\ell} \equiv T_{\v v_i\ell}$.
 The state $\rho$ can be expanded in the basis of these operators as
@@
 \end{align}
-where $\tilde{\v q}_{\ell i} = (q_{\ell,-\ell,i},q_{\ell,-\ell+1,i},\cdots q_{\ell,\ell,i})$, and we used the fact that $\bbk{\epsilon_{i\ell}\epsilon_{i'\ell}} = \delta_{ii'}\times\cov_\rho\p{T_{i\ell},T_{i\ell}}/n$.
+where $\tilde{\v q}_{\ell i} = \sum_m q_{\ell mi} \ket{m}$, and we used the fact that $\bbk{\epsilon_{i\ell}\epsilon_{i'\ell}} = \delta_{ii'}\times\cov_\rho\p{T_{i\ell},T_{i\ell}}/n$.
 Identifying the singular value decomposition $M_{V\ell} = U_{V\ell} \Sigma_{V\ell} W_{V\ell}^\dag$, we then we observe that vector $\tilde{\v q}_{\ell i}$ is essentially the $i$-th row of $\Sigma_{V\ell}^{-1} W_{V\ell}^\dag$, which is to say that $\tilde{\v q}_{\ell i} = \Sigma_{V\ell}^{-1} W_{V\ell}^\dag \ket{\v v_i}$.

### 2020-11-24 sun_phases/spin_qudit_tomo.tex
remove sub-section numbering in appendix

@@
 Though we cannot minimize over all suitable choices of measurement axes $V$ to compute $\beta(p)$, we can compute an empirical upper bound $\tilde\beta(p)\ge\beta(p)$ by minimizing over a large number of randomly chosen $V$.
-Figure \ref{fig:axes} shows the results of such empirical minimization, where we find that $\tilde\beta(p)$ drops substantially before plateauing at $p\approx d$, after which there are only marginal benefits to using more measurement axes.
+Figure \ref{fig:axes} shows the results of such empirical minimization, where we find that $\tilde\beta(p)$ drops substantially before plateauing at $p\approx d$, after which there are only minor benefits to using more measurement axes.
 In the interest of reducing experimental complexity as well as the runtime of our tomography protocol, which grow linearly in $\abs{V}$, we therefore conclude that spin qudit tomography should be performed with $\abs{V}\approx3d$ measurement axes.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Comments on a tight reconstruction error bound}
-\label{sec:tight}
+\subsection*{Comments on a tight reconstruction error bound}
 

### 2020-11-30 sun_phases/spin_qudit_tomo.tex
update to spin qudit tomography paper in response to feeback from AMR

@@
 
-Due to advancements in experimental capabilities to address nuclear spins in ultracold atomic systems \cite{cazalilla2014ultracold} [\red{more references?}], as well as developments in the control of ultracold molecular systems \cite{zeppenfeld2012sisyphus, kozyryev2017sisyphus, puri2017synthesis, wu2017cryofuge, marco2019degenerate, liu2019molecular, anderegg2019optical, chou2020frequencycomb, lin2020quantum, covey2018approach, jamadagni2019quantum}, a particular setting of growing interest is the spin qudit, or a multilevel quantum angular momentum degree of freedom.
-Spin qudits can provide advantages over their qubit counterparts for quantum sensing \cite{evrard2019enhanced}, enable quantum simulations of SU($d$) magnetism \cite{cazalilla2014ultracold, banerjee2013atomic, zhang2014spectroscopic, scazza2014observation, goban2018emergence}, and offer unique capabilities for quantum error correction \cite{albert2020robust, gross2020encoding}.
+Due to advancements in experimental capabilities to address nuclear spin states (hyperfine levels) in ultracold atomic systems \cite{daley2011quantum, smith2013quantum, cazalilla2014ultracold}, as well as developments in the control of ultracold molecular systems \cite{bohn2017cold, zeppenfeld2012sisyphus, kozyryev2017sisyphus, puri2017synthesis, wu2017cryofuge, marco2019degenerate, liu2019molecular, anderegg2019optical, chou2020frequencycomb, lin2020quantum}, a particular setting of growing interest is the spin qudit, or a multilevel quantum angular momentum degree of freedom.
+Spin qudits can provide advantages over their qubit counterparts for quantum sensing \cite{hemmer2018squeezing, evrard2019enhanced}, enable quantum simulations of SU($d$) magnetism \cite{cazalilla2014ultracold, banerjee2013atomic, zhang2014spectroscopic, scazza2014observation, goban2018emergence}, and offer unique capabilities for quantum error correction \cite{albert2020robust, gross2020encoding}.
 In all cases, quantum state tomography is necessary to take full advantage of a spin qudit.
@@
 For ease of notation, we will therefore implicitly associate Euler angle doublets $\v v=\p{\alpha,\beta}$ with the triplet $\p{\alpha,\beta,0}$, such that $R\p{\v v} \equiv R\p{\alpha,\beta,0}$, $D^\ell_{mn}\p{\v v} \equiv D^\ell_{mn}\p{\alpha,\beta,0}$, and $T_{\v v\ell m} \equiv T_{\p{\alpha,\beta,0},\ell m}$.
-We will also frequently consider polarization operators with order $m=0$.
-Symbols with a suppressed order therefore implicitly set this order to zero, which is to say that $T_\ell\equiv T_{\ell,0}$, $T_{\v v\ell}\equiv T_{\v v\ell,0}$, and $D^\ell_m\equiv D^\ell_{m,0}$.
 
 As a final point, we mention that the polarization operators $T_{\ell m}$ share a connection to the spherical harmonics $Y_{\ell m}$ that goes beyond the rules for their transformation under 3D rotations.
-In fact, the phase-space (Wigner function) representation of $T_{\ell m}$ is essentially equal to $Y_{\ell m}$ (up to complex conjugation and a normalization factor).
+In fact, the phase-space (Wigner function) representation of $T_{\ell m}$ is essentially equal to $Y_{\ell m}$, up to complex conjugation and a normalization factor.
 The phase-space representation of an operator $X$ assigns, to each point $\v v$ on the sphere, the complex number
@@
 This representation is faithful in the sense that $X$ is uniquely determined by the phase-space values $\tilde X\p{\v v}$ at all points $\v v$ on the sphere.
-The transformation rules for polarization operators in Eq.~\eqref{eq:trans_rot}, together with the fact that $\bk{s|T_{\ell m}|s}=0$ unless $m=0$, suffice to show that their phase-space distributions are
+The transformation rules for polarization operators in Eq.~\eqref{eq:trans_rot}, together with the fact that $\bk{s|T_{\ell m}|s}=0$ unless $m=0$, suffice to show that their phase-space distributions are (see Appendix \ref{sec:rotations})
 \begin{align}
@@
 \end{align}
-where the proportionality factor $c_\ell$ simply enforces that $\obk{T_{\ell m}|T_{\ell m}}=1$ (see Appendix \ref{sec:rotations}).
+where the scalar $c_\ell$ simply enforces that $\obk{T_{\ell m}|T_{\ell m}}=1$.
 In this way, the polarization operators $T_{\ell m}$ are a quantum analogue of spherical harmonics $Y_{\ell m}$, which play an important role in phase-space formalisms for spin qudits \cite{li2013weylwignermoyal}.
@@
 We are thus nominally restricted to measuring projectors $\Pi_{\v v\mu} \equiv \op{\mu_{\v v}}$, where $\ket{\mu_{\v v}}\equiv R\p{\v v}\ket{\mu}$ is a state with spin projection $\mu$ onto the measurement axis $\v v$.
-For any fixed axis $\v v$, the sets $\set{\Pi_{\v v\mu}}$ and $\set{T_{\v v\ell}}$ are both complete bases for the space of operators that are diagonal in the basis $\set{\ket{\mu_{\v v}}}$.
-Measuring the projectors $\set{\Pi_{\v v\mu}}$ is therefore equivalent to measuring the polarization operators $\set{T_{\v v\ell}}$, and provides data on the expectation values $\bk{T_{\v v\ell}}_\rho$.
-
-In order to reconstruct an arbitrary density operator $\rho$ from the expectation values $\bk{T_{\v v\ell}}_\rho$, we essentially need to to find a set of coefficients $C_{\ell mk}\p{\v v}$ that would allow us to recover any matrix element $\rho_{\ell m}$ of $\rho$ through
-\begin{align}
-  \rho_{\ell m} = \bk{T_{\ell m}^\dag}_\rho
-  = \sum_{\v v,k} C_{\ell mk}\p{\v v} \bk{T_{\v vk}^\dag}_\rho.
+For any fixed axis $\v v$, the sets $\set{\Pi_{\v v\mu}}$ and $\set{T_{\v v\ell,0}}$ are both complete bases for the space of operators that are diagonal in the basis $\set{\ket{\mu_{\v v}}}$.
+Measuring the projectors $\set{\Pi_{\v v\mu}}$ is therefore equivalent to measuring the polarization operators $\set{T_{\v v\ell,0}}$, and provides data on the expectation values $\bk{T_{\v v\ell,0}}_\rho$.
+
+In order to reconstruct an arbitrary density operator $\rho$ from the expectation values $\bk{T_{\v v\ell,0}}_\rho$, we essentially need to to find a set of coefficients $C_{\ell mk}\p{\v v}$ that would allow us to recover any matrix element $\rho_{\ell m}$ of $\rho$ through
+\begin{align}
+  \rho_{\ell m}^* = \bk{T_{\ell m}}_\rho
+  = \sum_{\v v,k} C_{\ell mk}\p{\v v} \bk{T_{\v vk,0}}_\rho.
   \label{eq:state_recon}
 \end{align}
-Expanding the rotated polarization operators $T_{\v vk}$ into a sum of un-rotated polarization operators, according to Eq.~\eqref{eq:trans_rot}, we find that the recovery condition in Eq.~\eqref{eq:state_recon} is satisfied when
-\begin{align}
-  T_{\ell m}^\dag
+Expanding the rotated polarization operators $T_{\v vk,0}$ into a sum of un-rotated polarization operators, according to Eq.~\eqref{eq:trans_rot}, we find that the recovery condition in Eq.~\eqref{eq:state_recon} is satisfied when
+\begin{align}
+  T_{\ell m}
   = \sum_{\v v,k,n} C_{\ell mk}\p{\v v}
-  D^k_n\p{\v v}^* T_{kn}^\dag.
+  D^k_{n,0}\p{\v v} T_{kn}.
 \end{align}
@@
 \begin{align}
-  \sum_{\v v} C_{\ell m}\p{\v v} D^\ell_n\p{\v v}^* = \delta_{mn}
+  \sum_{\v v} C_{\ell m}\p{\v v} D^\ell_{n,0}\p{\v v} = \delta_{mn}
   \label{eq:tomo_recovery}
@@
 We will refer to Eq.~\eqref{eq:full_recovery} as the {\it full recovery problem}, and Eq.~\eqref{eq:tomo_recovery} as the the {\it reduced recovery problem}.
-Due to the fact that $D^k_n\p{\v v}^* = \sqrt{\frac{4\pi}{2\ell+1}}\, Y_{kn}\p{\v v}$, any solution to the full recovery problem automatically solves the reduced recovery problem by setting $C_{\ell m}\p{\v v} = \sqrt{\frac{2\ell+1}{4\pi}}\, \tilde C_{\ell m}\p{\v v}$.
+Due to the fact that $D^k_{n,0}\p{\v v} = \sqrt{\frac{4\pi}{2\ell+1}}\, Y_{kn}\p{\v v}^*$, any solution to the full recovery problem automatically solves the reduced recovery problem by setting $C_{\ell m}\p{\v v} = \sqrt{\frac{2\ell+1}{4\pi}}\, \tilde C_{\ell m}\p{\v v}^*$.
 In principle, this mapping allows us to import a host of existing signal recovery algorithms \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality} for the task of spin qudit tomography.
@@
 \begin{enumerate*}
-\item there are only $2\ell+1\le2d-1$ polarization operators of degree $\ell\le\ell_{\t{max}}$, and
-\item measuring spin projection along any single axis provides measurement data for polarization operators of {\it all} degrees $\ell\le\ell_{\t{max}}$,
+\item there are at most $2d-1$ polarization operators of any given degree, and
+\item measuring spin projection along any single axis provides measurement data for polarization operators of all degrees,
 \end{enumerate*}
 one might expect that $2d-1$ suitably chosen measurement axes would always suffice to perform full tomography of a $d$-level spin qudit.
-In the language of classical signal recovery, we would expect there to exist a set of $2\ell_{\t{max}}+1$ points on the sphere that enable the recovery of {\it any} function with fixed degree $\ell\le\ell_{\t{max}}$, given prior knowledge of $\ell$ \footnote{We could further conjecture that if {\it any} set of points $V\subset\Omega_2$ enables the recovery of degree-$\ell_{\t{max}}$ functions on the sphere, then $V$ enables the recovery degree-$\ell$ functions with $\ell\le\ell_{\t{max}}$.
+In the language of classical signal recovery, we would expect there to exist a set of $2\ell_{\t{max}}+1$ points on the sphere that enable the recovery of any function with fixed degree $\ell\le\ell_{\t{max}}$, given prior knowledge of $\ell$ \footnote{We could further conjecture that if {\it any} set of points $V\subset\Omega_2$ enables the recovery of degree-$\ell_{\t{max}}$ functions on the sphere, then $V$ enables the recovery degree-$\ell$ functions with $\ell\le\ell_{\t{max}}$.
   However, this condition is stronger than necessary to guarantee full recovery of a $d$-level spin qudit state with $2d-1$ measurement axes, which only requires the existence of {\it some} such set of points.}.
@@
 A set of measurement axes $V=\set{\v v}$ nominally induces a set of projectors $\set{\Pi_{\v v\mu}}$ that will be measured in an experiment.
-By a simple change of basis, measuring these projectors is equivalent to measuring the polarization operators $\set{T_{\v v\ell}}$.
-Flattening each $d\times d$ matrix $T_{\v v\ell}$ into the $d^2$-component column vector $\oket{T_{\v v\ell}}$, we construct the {\it measurement matrix}
-\begin{align}
-  M_V \equiv \sum_{\v v,\ell} \oket{T_{\v v\ell}} \bra{\v v\ell},
+By a simple change of basis, measuring these projectors is equivalent to measuring the polarization operators $\set{T_{\v v\ell,0}}$.
+Flattening each $d\times d$ matrix $T_{\v v\ell,0}$ into the $d^2$-component column vector $\oket{T_{\v v\ell,0}}$, we construct the {\it measurement matrix}
+\begin{align}
+  M_V \equiv \sum_{\v v,\ell} \oket{T_{\v v\ell,0}} \bra{\v v\ell}.
   \label{eq:meas_mat}
@@
 Here $\v v$ and $\ell$ label a column of $M_V$, or equivalently label a standard basis vector $\ket{\v v\ell}$ of a $\p{\abs{V}\times d}$-dimensional vector space.
-A necessary and sufficient condition for $V$ to allow for full state tomography is that the measured polarization operators in $\set{T_{\v v\ell}}$, or equivalently the columns of $M_V$, span the entire ($d^2$-dimensional) space of operators on a $d$-level spin qudit.
+A necessary and sufficient condition for $V$ to allow for full state tomography is that the measured polarization operators in $\set{T_{\v v\ell,0}}$, or equivalently the columns of $M_V$, span the entire ($d^2$-dimensional) space of operators on a $d$-level spin qudit.
 In this case $M_V$ must be full rank, with $d^2$ nonzero singular values.
@@
 \end{align}
-where for shorthand we use a combined index $j=\p{\v v,\ell}$ to specify both a measurement axis $\v v$ and a degree $\ell$, which identify the polarization operator $T_j\equiv T_{\v v\ell}$.
+where for shorthand we use a combined index $j=\p{\v v,\ell}$ to specify both a measurement axis $\v v$ and a degree $\ell$, which identify the polarization operator $T_j\equiv T_{\v v\ell,0}$.
 These operators allow us to expand any state $\rho$ of a $d$-level spin qudit in the form
@@
 \begin{align}
-  M_{V\ell} \equiv \sum_{m,\v v} \ket{m} \obk{T_{\ell m}|T_{\v v\ell}} \bra{\v v}
-  = \sum_{m,\v v} D^\ell_m\p{\v v} \op{m}{\v v}.
+  M_{V\ell} \equiv \sum_{m,\v v} \ket{m} \obk{T_{\ell m}|T_{\v v\ell,0}} \bra{\v v}
+  = \sum_{m,\v v} D^\ell_{m,0}\p{\v v} \op{m}{\v v}.
   \label{eq:block}
@@
     Each point is an average over $10^3$ calculations or 5 minutes of runtime, whichever comes first.
-    These results do not count fixed runtimes to pre-compute quantities that can be recycled for every new choice of $V$ or $\rho$.
-    Dashed lines show fits to a runtime $t=c d^\alpha$ for all $d\ge100$, finding $\alpha\approx 3.7$.
+    These results do not count fixed runtimes to pre-compute quantities that can be recycled for every new choice of $V$ and $\rho$.
+    Dashed lines show fits to a runtime $t=c d^\alpha$ for the 20 largest values of $d$, finding $\alpha\approx 3.8\pm0.1$.
     These simulations found axes $V$ with $\epsilon_V<d$ for all $d\le200$, thereby verifying for these cases that $2d-1$ measurement axes are sufficient to perform full state tomography.}
@@
 \end{align}
-where the {\it quantum error scale} $\epsilon_V$, so called because it is set by quantum shot noise, is defined in terms of (half of) the spectral range of $T_\ell$:
+where the {\it quantum error scale} $\epsilon_V$, so called because the bound in Eq.~\eqref{eq:bound} is set by quantum shot noise, is defined in terms of (half of) the spectral range of $T_{\ell,0}$:
 \begin{align}
@@
   &&
-  t_{\ell\mu} \equiv \bk{\mu|T_\ell|\mu}.
+  t_{\ell\mu} \equiv \bk{\mu|T_{\ell,0}|\mu}.
 \end{align}
@@
 
-Though straightforward to compute, the bound in Eq.~\eqref{eq:bound} is not tight, as it is acquired by bounding the statistical error in the empirical estimate $\tilde T_{\v v\ell}$ of $\bk{T_{\v v\ell}}_\rho$ by $\bbk{\epsilon_{\v v\ell}^2}\le\Gamma_\ell^2$.
-Though the individual bounds on $\bbk{\epsilon_{\v v\ell}^2}$ for each axis $\v v$ and degree $\ell$ are tight, these bounds cannot all be achieved simultaneously.
+Though straightforward to compute, the bound in Eq.~\eqref{eq:bound} is not tight, as it is acquired by bounding the statistical error in the empirical estimate $\tilde T_{\v v\ell,0}$ of $\bk{T_{\v v\ell,0}}_\rho$ by $\bbk{\epsilon_{\v v\ell,0}^2}\le\Gamma_\ell^2$.
+The individual bounds on $\bbk{\epsilon_{\v v\ell,0}^2}$ for each axis $\v v$ and degree $\ell$ are tight, these bounds cannot all be achieved simultaneously.
 There is therefore still room for improvement on the bound in Eq.~\eqref{eq:bound} by maximizing $\E_V$ over the set of all qudit states $\rho$.
@@
 \end{align}
-where $\ket{\rho_\ell} \equiv \sum_m \rho_{\ell m}\ket{m}$, and $\ket{\chi_{V\ell}}$ and $\N_{V\ell}$ are defined below.
-While the true error in $\tilde\rho_V$ cannot be known exactly without knowing $\rho$, this error can be estimated by $\E_V\p{\tilde\rho_V} \approx \E_V\p{\rho}$.
-
-The matrix $\N_{V\ell}$ that appears in Eq.~\eqref{eq:error} is defined in terms of the measurement matrix $M_{V\ell}$ and its right inverse $M_{V\ell}^{-1}$ by
+where $\ket{\rho_\ell} \equiv \sum_m \rho_{\ell m}\ket{m}$ is a vector of the degree-$\ell$ polarization operator components $\rho_{\ell m}$ of $\rho$, defined in Eq.~\eqref{eq:trans_state}, and the matrix $\N_{V\ell}$ and vector $\ket{\chi_{V\ell}}$ are defined below.
+While the true error in $\tilde\rho_V$ cannot be known exactly without knowing $\rho$, this error can be estimated a posteriori by $\E_V\p{\tilde\rho_V} \approx \E_V\p{\rho}$.
+After constructing an estimate $\tilde\rho_V$ of $\rho$, the complexity of computing $\E_V\p{\tilde\rho_V}$ is the same as that of $\S_V$ and $\epsilon_V$ (see Figure \ref{fig:times}).
+
+We now define $\N_{V\ell}$ and $\ket{\chi_{V\ell}}$ for the sake of completion, but note that these definitions can be skipped without consequence for the remaining discussions in this paper.
+The matrix $\N_{V\ell}$ is defined in terms of the measurement matrix $M_{V\ell}$ and its right inverse $M_{V\ell}^{-1}$ by
 \begin{align}
@@
 where $\diag\sp{X}$ sets all off-diagonal parts of $X$ to zero.
-The vector $\ket{\chi_{V\ell}}\equiv\sum_m \chi^V_{\ell m} \ket{m}$ in Eq.~\eqref{eq:error} is defined by
+The vector $\ket{\chi_{V\ell}}\equiv\sum_m \chi^V_{\ell m} \ket{m}$ is defined by
 \begin{align}
@@
 \end{align}
-Here $\tilde g_{L\ell}$ is essentially a matrix of structure constants for the polarization operator algebra (see Appendix \ref{sec:trans_prod}), and $\D_M$ simply picks off the $M$-th diagonal of the matrix it acts on, such that $\obk{\N_{V\ell}|\D_M|\tilde g_{L\ell}}$ is an inner product of the $M$-th diagonal of $\tilde g_{L\ell}$ with the $\p{-M}$-th diagonal of $\N_{V\ell}$.
-Overall, the complexity of computing $\E_V\p{\rho}$ is the same as that of $\S_V$ and $\epsilon_V$ (see Figure \ref{fig:times}).
+Here $\tilde g_{L\ell}$ is essentially a matrix of structure constants for the polarization operator algebra (see Appendix \ref{sec:trans_prod}), and $\D_M$ simply picks off the $M$-th diagonal of the matrix it acts on.
 
@@
 \begin{align}
-  \tilde T_{\v v\ell}
-  \equiv \sum_\mu \bk{\mu|T_\ell|\mu} \tilde\Pi_{\v v\mu}
-\end{align}
-of $\bk{T_{\v v\ell}}_\rho$, which can in turn be used to construct the estimate $\tilde\rho_V$ of $\rho$.
+  \tilde T_{\v v\ell,0}
+  \equiv \sum_\mu \bk{\mu|T_{\ell,0}|\mu} \tilde\Pi_{\v v\mu}
+\end{align}
+of $\bk{T_{\v v\ell,0}}_\rho$, which can in turn be used to construct the estimate $\tilde\rho_V$ of $\rho$.
 Specifically, the $2\ell+1$ (normalized) right singular vectors $\v x^V_{\ell k} = \sum_{\v v} x^V_{\ell k \v v} \ket{\v v}$ and corresponding singular values $M^V_{\ell k}$ of the matrices $M_{V\ell}$ defined in Eq.~\eqref{eq:block} determine the components
@@
   = \f1{M^V_{\ell k}} \sum_{\v v}
-  x^V_{\ell k\v v} \, \tilde T_{\v v\ell},
+  x^V_{\ell k\v v} \, \tilde T_{\v v\ell,0},
 \end{align}
@@
   Q_{\ell k} = \f1{M^V_{\ell k}} \sum_{\v v}
-  x^V_{\ell k\v v} T_{\v v\ell}.
+  x^V_{\ell k\v v} T_{\v v\ell,0}.
 \end{align}
@@
 Though $2d-1$ measurement axes may be sufficient to perform full state tomography, this is not necessarily the best choice of $\abs{V}$.
-Increasing the number of measurement axes generally decreases the quantum error scale $\epsilon_V$, but at the cost of having to estimate more observables.
+Increasing the number of measurement axes generally decreases the quantum error scale $\epsilon_V$, but comes at the cost of having to estimate more observables.
 At a fixed total number of measurements, increasing $\abs{V}$ reduces the number of measurements $n$ devoted to each axis $\v v\in V$.
@@
 
-Rather than the statistical errors $\epsilon_{i\mu} \equiv \tilde\Pi_{i\mu} - \bk{\Pi_{i\mu}}_\rho$ in the estimates $\tilde\Pi_{i\mu}$ of the projectors $\Pi_{i\mu}$, we now consider the statistical errors $\epsilon_{i\ell} \equiv \tilde T_{i\ell} - \bk{T_{i\ell}}_\rho$ in the estimates $\tilde T_{i\ell}$ of the polarization operators $T_{i\ell} \equiv T_{\v v_i\ell}$.
+Rather than the statistical errors $\epsilon_{i\mu} \equiv \tilde\Pi_{i\mu} - \bk{\Pi_{i\mu}}_\rho$ in the estimates $\tilde\Pi_{i\mu}$ of the projectors $\Pi_{i\mu}$, we now consider the statistical errors $\epsilon_{i\ell} \equiv \tilde T_{i\ell} - \bk{T_{i\ell}}_\rho$ in the estimates $\tilde T_{i\ell}$ of the polarization operators $T_{i\ell} \equiv T_{\v v_i\ell,0}$.
 We can expand the polarization operators $T_{i\ell}$ as a sum over projectors $\Pi_{i\mu}$ as
@@
   &&
-  t_{\ell\mu} \equiv \bk{\mu|T_\ell|\mu}
+  t_{\ell\mu} \equiv \bk{\mu|T_{\ell,0}|\mu}
   = \sqrt{\f{2\ell+1}{d}} \bk{s\mu;\ell|s\mu},
@@
 
-To this end, we define the probability $p^i_\mu\equiv\bk{\Pi_{i\mu}}_\rho$, collect these probabilities into the classical probability distribution $p^i = \sum_\mu p^i_\mu \ket{\mu}$, and define the vector $t_\ell \equiv \sum_\mu t_{\ell\mu} \ket{\mu}$ of all $t_{\ell\mu}\equiv\bk{\mu|T_\ell|\mu}$.
+To this end, we define the probability $p^i_\mu\equiv\bk{\Pi_{i\mu}}_\rho$, collect these probabilities into the classical probability distribution $p^i = \sum_\mu p^i_\mu \ket{\mu}$, and define the vector $t_\ell \equiv \sum_\mu t_{\ell\mu} \ket{\mu}$ of all $t_{\ell\mu}\equiv\bk{\mu|T_{\ell,0}|\mu}$.
 We then observe that
@@
 \begin{align}
-  M_V \equiv \sum_{\v v,\ell} \oket{T_{\v v\ell}} \bra{\v v\ell},
+  M_V \equiv \sum_{\v v,\ell} \oket{T_{\v v\ell,0}} \bra{\v v\ell},
 \end{align}
@@
   &&
-  M_{V\ell} = \sum_{m,\v v} D^\ell_m\p{\v v} \op{m}{\v v},
+  M_{V\ell} = \sum_{m,\v v} D^\ell_{m,0}\p{\v v} \op{m}{\v v},
 \end{align}
@@
   = \cov_\rho\p{T_{i\ell}^\dag,T_{i\ell}}
-  = \sum_{m,m'} D^\ell_m\p{\v v_i}^* D^\ell_{m'}\p{\v v_i}
+  = \sum_{m,m'} D^\ell_{m,0}\p{\v v_i}^* D^\ell_{m',0}\p{\v v_i}
   \cov_\rho\p{T_{\ell m}^\dag, T_{\ell m'}},
@@
   = \f1n \sum_{i,\ell,m,m'}
-  D^\ell_{m'}\p{\v v_i} \abs{\tilde{\v q}_{\ell i}}^2
-  D^\ell_m\p{\v v_i}^*
+  D^\ell_{m',0}\p{\v v_i} \abs{\tilde{\v q}_{\ell i}}^2
+  D^\ell_{m,0}\p{\v v_i}^*
   \cov_\rho\p{T_{\ell m}^\dag, T_{\ell m'}}.

### 2020-11-30 sun_phases/spin_qudit_tomo.tex
fixed minor error concerning the phase-spase representation of polarization operators

@@
 As a final point, we mention that the polarization operators $T_{\ell m}$ share a connection to the spherical harmonics $Y_{\ell m}$ that goes beyond the rules for their transformation under 3D rotations.
-In fact, the phase-space (Wigner function) representation of $T_{\ell m}$ is essentially equal to $Y_{\ell m}$, up to complex conjugation and a normalization factor.
-The phase-space representation of an operator $X$ assigns, to each point $\v v$ on the sphere, the complex number
-\begin{align}
-  \tilde X\p{\v v} \equiv \bk{s_{\v v}|X|s_{\v v}},
+In fact, the phase-space (Wigner function) representation of $T_{\ell m}$ is proportional to $Y_{\ell m}$.
+The phase-space representation of a spin qudit operator $X$ assigns, to each point $\v v$ on the sphere, the number
+\begin{align}
+  X^{\t{PS}}\p{\v v} \equiv \bk{s_{\v v}|X|s_{\v v}},
 \end{align}
 where $\ket{s_{\v v}}\equiv R\p{\v v}\ket{s}$ is the state of a spin qudit polarized along $\v v$.
-This representation is faithful in the sense that $X$ is uniquely determined by the phase-space values $\tilde X\p{\v v}$ at all points $\v v$ on the sphere.
-The transformation rules for polarization operators in Eq.~\eqref{eq:trans_rot}, together with the fact that $\bk{s|T_{\ell m}|s}=0$ unless $m=0$, suffice to show that their phase-space distributions are (see Appendix \ref{sec:rotations})
-\begin{align}
-  \tilde T_{\ell m}\p{\v v} = c_\ell Y_{\ell m}\p{\v v}^*,
+This representation is faithful in the sense that $X$ is uniquely determined by the phase-space values $X^{\t{PS}}\p{\v v}$ at all points $\v v$ on the sphere.
+The transformation rules for polarization operators in Eq.~\eqref{eq:trans_rot}, together with the fact that $\bk{s|T_{\ell m}|s}=0$ unless $m=0$, suffice to show that (see Appendix \ref{sec:rotations})
+\begin{align}
+  T_{\ell m}^{\t{PS}}\p{\v v} = c_\ell Y_{\ell m}\p{\v v},
 \end{align}
 where the scalar $c_\ell$ simply enforces that $\obk{T_{\ell m}|T_{\ell m}}=1$.
-In this way, the polarization operators $T_{\ell m}$ are a quantum analogue of spherical harmonics $Y_{\ell m}$, which play an important role in phase-space formalisms for spin qudits \cite{li2013weylwignermoyal}.
+The polarization operators $T_{\ell m}$ are a quantum analogue of spherical harmonics $Y_{\ell m}$, which play an important role in phase-space formalisms for spin qudits \cite{li2013weylwignermoyal}.
 
@@
 \begin{align}
-  D^\ell_{0,m}\p{\v v} = \p{-1}^m D^\ell_{-m,0}\p{\v v}^*
-  = \sqrt{\f{4\pi}{2\ell+1}} \p{-1}^m Y_{\ell,-m}\p{\v v}
-  = \sqrt{\f{4\pi}{2\ell+1}} \, Y_{\ell m}\p{\v v}^*,
+  D^\ell_{0,m}\p{\v v} = \sqrt{\f{4\pi}{2\ell+1}} \, Y_{\ell m}\p{\v v},
 \end{align}
@@
     \p{\f{\p{2s}!}{\p{2s+\ell}!}} \p{\f{\p{2s}!}{\p{2s-\ell}!}}} \,
-  Y_{\ell m}\p{\v v}^*
-  \propto Y_{\ell m}\p{\v v}^*.
+  Y_{\ell m}\p{\v v}.
 \end{align}

### 2020-11-30 sun_phases/spin_qudit_tomo.tex
minor change to footnote, and added a missing word

@@
 where the scalar $c_\ell$ simply enforces that $\obk{T_{\ell m}|T_{\ell m}}=1$.
-The polarization operators $T_{\ell m}$ are a quantum analogue of spherical harmonics $Y_{\ell m}$, which play an important role in phase-space formalisms for spin qudits \cite{li2013weylwignermoyal}.
+The polarization operators $T_{\ell m}$ are thus a quantum analogue of spherical harmonics $Y_{\ell m}$, which play an important role in phase-space formalisms for spin qudits \cite{li2013weylwignermoyal}.
 
@@
 If the function $f$ is {\it band-limited} at degree $L$, which is to say that $f_{\ell m}=0$ for all $\ell\ge L$, then the full recovery problem in Eq.~\eqref{eq:full_recovery} is provably solvable with a suitable choice of $\abs{V}=L^2$ points on the sphere \cite{freeden2008spherical, freeden2018spherical}.
-The existence of these solutions to the full recovery problem in turn implies the existence of $d^2$ measurement axes that suffice to reconstruct arbitrary states of $d$-level spin qudit, whose possible states are essentially ``band-limited'' at degree $d$ \footnote{Spin qudit states can be faithfully represented by quasi-probability distributions on a sphere.
-  For a $d$-level spin, these quasi-probability distributions are band-limited at degree $d$.}.
+The existence of these solutions to the full recovery problem in turn implies the existence of $d^2$ measurement axes that suffice to reconstruct arbitrary states of $d$-level spin qudit, whose possible states are essentially band-limited at degree $d$ \footnote{Specifically, the phase-space representation of a $d$-level spin qudit state $\rho$ is band-limited at degree $d$.}.
 Moreover, for any fixed degree $\ell$, finding solutions to the reduced recovery problem in Eq.~\eqref{eq:tomo_recovery} is equivalent to the recovery of a degree-$\ell$ function $f_\ell = \sum_m f_{\ell m} Y_{\ell m}$, which is provably possible with $\abs{V}=2\ell+1$ samples \cite{freeden2008spherical}.

### 2020-12-01 sun_phases/sun_dynamics_notes.tex
fixed errors in the 'simple model' section

@@
 
-\subsection{Hamiltonian}
-
 Here we consider a simple Hamiltonian on $2N$ spins,
@@
 \begin{align}
-  H = U \v S_0 \c\v S_1 + S_{\z,0} - S_{\z,1}
-  = \sum_{m,n=-IN}^{IN} \p{U\op{mn}{nm}
-    + \p{m-n}\op{mn}},
+  H = U \v S_0 \c\v S_1 + S_{\z,0} - S_{\z,1}.
 \end{align}
-where $m,n$ are projections of a net sub-ensemble spin onto the $z$ axis, and we have neglected the scalars $\v S_k\cdot \v S_k = IN\p{IN+1}$.
-Indexing states $\ket{mn}$ by the mean spin projection $\bar m=\p{m+n}/2$ and spin imbalance $\delta=m-n$,
-\begin{align}
-  \ket{\bar m\delta} \equiv \ket{m+\delta/2,m-\delta/2},
-\end{align}
-we can expand
-\begin{align}
-  \v S_0\c \v S_1 = \sum_{m,\delta}
-  \op{\bar m,-\delta}{\bar m\delta},
-  &&
-  S_{\z,0} - S_{\z,1} = \sum_{m,\delta} \delta \op{\bar m\delta},
-\end{align}
-which allows us to write the Hamiltonian in the form
-\begin{align}
-  H = \sum_m \op{\bar m} \otimes
-  \sum_\delta \p{U \op{-\delta}{\delta} + \delta \op{\delta}}.
-\end{align}
-Diagonalizing this Hamiltonian thus reduces to a qubit problem, for which we define the Pauli operator
-\begin{align}
-  X_\delta\p{U} \equiv \sum_{\sigma\in\set{\pm1}}
-  \sigma \op{\delta,U,\sigma},
-\end{align}
-in terms of the rotated states
-\begin{align}
-  \ket{\delta,U,\sigma} \equiv \cos\p{\f{\eta}{2}} \ket{\p{\sigma\delta}}
-  + \sigma \sin\p{\f{\eta}{2}} \ket{\p{-\sigma\delta}},
-  &&
-  \eta \equiv \arctan\p{\f{U}{\delta}},
-\end{align}
-defined for $\delta\ne0$.
-In words, the state $\ket{\delta,U,\pm}$ points in the direction of $\pm\p{\delta\uv{z}+U\uv{x}}$ on the Bloch sphere defined by the state $\set{\ket{\delta},\ket{-\delta}}$.
-Altogether, the Hamiltonian is
-\begin{align}
-  H = \sum_m \op{\bar m} \otimes
-  \sp{U\op{0} + \sum_{\delta>0} U_\delta X_\delta\p{U}},
-  &&
-  U_\delta \equiv \sqrt{U^2+\delta^2},
-\end{align}
-which has eigenvalues $U$ and $\pm U_\delta$ for integer $\delta>0$.
 
-\subsection{Initial state}
+\subsection{Simplifying the Hamiltonian}
 
-We now consider an initial state with both large spins polarized at a polar angle $\theta$ and azimuthal angles $\pm\phi/2$:
-\begin{align}
-  \ket{\theta,\phi} \equiv \sum_{m,n}
-  \sqrt{\chi_m\p{\theta} \chi_n\p{\theta}} \,
-  e^{-\i\p{m-n}\phi} \ket{mn}
-  = \sum_{m,\delta}
-  \chi_{m\delta}\p{\theta} e^{-\i\delta\phi}
-  \ket{\bar m\delta}.
-\end{align}
-where
-\begin{align}
-  \chi_m\p\theta \equiv { 2IN \choose IN + m }
-  \cos\p{\f{\theta}{2}}^{2IN+2m} \sin\p{\f{\theta}{2}}^{2IN-2m},
-  &&
-  \chi_{m\delta}\p\theta
-  \equiv \sqrt{\chi_{m+\delta/2}\p\theta \chi_{m-\delta/2}\p\theta}.
-\end{align}
-In the eigenbasis of $H$, we have
-\begin{align}
-  \ket{\theta,\phi} = \sum_m \ket{\bar m} \otimes
-  \sp{\chi_0\p\theta \ket{0}
-    + \sum_{\delta>0} \chi_{m\delta}\p\theta
-    \sum_{\sigma\in\set{\pm1}} e^{-\i\sigma\delta\phi}
-    \sp{\cos\p{\f{\eta}{2}} \ket{\delta,U,\sigma}
-    - \sigma\sin\p{\f{\eta}{2}} \ket{\delta,U,-\sigma}}},
-\end{align}
-where again $\eta\equiv\arctan\p{U/\delta}$ for shorthand.
-Note that the only dependence of this state on the nuclear spin $I$ appears in the range of $\bar m$, so any distinction between observed behavior at different $I$ can essentially be attributed to the measurement of different observables.
-
-\subsection{Observables}
-
-We now turn our attention to the measurement of collective observables.
-In particular, we consider the collective $N$-spin operators
+To simplify the Hamiltonian, we consider the collective $N$-spin operators
 \begin{align}
@@
   \tilde S_{\mu\epsilon} \equiv S_{\mu+\epsilon,\mu}
-  = \sum_m \xi_{\mu\epsilon}^m \op{m+\epsilon}{m},
+  = \sum_m c_{\mu\epsilon}^m \op{m+\epsilon}{m},
   &&
-  \xi_{\mu\epsilon}^m \equiv \bk{m+\epsilon|S_{\mu+\epsilon,\mu}|m}.
+  c_{\mu\epsilon}^m \equiv \bk{m+\epsilon|S_{\mu+\epsilon,\mu}|m}.
 \end{align}
-In turn, we symmetrize these operators over their action on two sub-ensembles of $N$ spins each:
+We can thus expand
 \begin{align}
-  \tilde S_{\mu\epsilon}^\sigma \equiv \tilde S_{\mu\epsilon} \otimes \1
-  + \sigma \1 \otimes \tilde S_{\mu\epsilon}
-  = \sum_{m,n} \p{\xi_{\mu\epsilon}^m \ket{m+\epsilon,n}
-    + \sigma \xi_{\mu\epsilon}^n \ket{m,n+\epsilon}} \bra{mn},
+  \v S_0 \c \v S_1
+  = \sum_{\mu,\epsilon} \tilde S_{\mu\epsilon,0}
+  \tilde S_{\mu\epsilon,1}^\dag
+  = \sum_{m,n,\epsilon} \xi^\epsilon_{m,n}
+  \op{m+\epsilon,n}{m,n+\epsilon},
+  &&
+  \xi^\epsilon_{m,n} \equiv
+  \sum_\mu c^m_{\mu\epsilon} \p{c^n_{\mu\epsilon}}^*.
 \end{align}
-where $\sigma\in\set{\pm1}$.
-In the symmetrized basis of a mean spin and spin imbalance, we have that
+Indexing states $\ket{mn}$ by the mean spin projection $M=\p{m+n}/2$ and spin imbalance $\delta=m-n$,
 \begin{align}
-  \tilde S_{\mu\epsilon}^\sigma
-  = \sum_{m,\delta,\sigma} \p{\xi_{\mu\epsilon}^{m+\delta/2}
-    \ket{\bar m_\epsilon,\delta+\epsilon}
-    + \sigma \xi_{\mu\epsilon}^{m-\delta/2}
-    \ket{\bar m_\epsilon,\delta-\epsilon}}
-  \bra{\bar m\delta},
+  \ket{M\delta} \leftrightarrow \ket{m+\delta/2,n-\delta/2},
+\end{align}
+we can expand
+\begin{align}
+  U \v S_0 \c \v S_1
+  = \sum_{M,\delta,\epsilon} U_{M\delta}^\epsilon
+  \op{M,\delta+\epsilon}{M,\delta-\epsilon},
   &&
-  \bar m_\epsilon \equiv \bar m + \epsilon/2.
+  U_{M\delta}^\epsilon \equiv
+  U \xi^\epsilon_{M+\delta/2-\epsilon/2,M-\delta/2-\epsilon/2}.
 \end{align}
-
-\newpage
+Additionally using the fact that
+\begin{align}
+  S_{\z,0} - S_{\z,1} = \sum_{M,\delta} \delta \op{M\delta},
+\end{align}
+the Hamiltonian is thus
+\begin{align}
+  H = \sum_M \op{M} \otimes
+  \sp{\sum_{\delta,\epsilon} U_{M\delta}^\epsilon \op{\delta+\epsilon}{\delta-\epsilon}
+    + \sum_\delta \delta \op\delta}.
+\end{align}
+For each value of $M$, this Hamiltonian looks like that of a single particle on a 1-D ``$\delta$-lattice'' with a linear tilt and position-dependent hopping.
+The parity of $\delta$ is conserved by this Hamiltonian, so in fact there is a single 1-D lattice for every value of $M$ and parity of $\delta$.
+The fact that $\epsilon\in\set{-2I,-2I+1,\cdots,2I}$ implies that the particle can hop as many as $2I$ sites on each fixed-parity lattice, with the hopping amplitude determined by $U_{M\delta}^\epsilon$.
 

### 2020-12-01 sun_phases/spin_qudit_tomo.tex
added reference, fixed typo, fixed bug with font encodings

@@
 }
+
+%%% properly deal with different font encodings
+\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
 
@@
 Quantum state tomography, or the task of reconstructing a quantum state by collecting and processing measurement data, is an essential primitive for quantum sensing, quantum simulation, and quantum information processing.
-The central importance of quantum state tomography has led to the development of a variety of techniques based on compressed sensing \cite{gross2010quantum}, matrix product states \cite{cramer2010efficient}, maximum-likelihood estimation \cite{smolin2012efficient}, Bayesian inference \cite{huszar2012adaptive}, and neural networks \cite{torlai2018neuralnetwork}, among many others \cite{christandl2012reliable, qi2013quantum, opatrny1997leastsquares}
+The central importance of quantum state tomography has led to the development of a variety of techniques based on compressed sensing \cite{gross2010quantum}, matrix product states \cite{cramer2010efficient}, maximum-likelihood estimation \cite{smolin2012efficient}, Bayesian inference \cite{huszar2012adaptive}, and neural networks \cite{torlai2018neuralnetwork}, among many others \cite{christandl2012reliable, qi2013quantum, opatrny1997leastsquares}.
 These techniques are typically developed in a general, information-theoretic setting, and make minimal assumptions about the physical medium of a quantum state.
@@
 
-Due to advancements in experimental capabilities to address nuclear spin states (hyperfine levels) in ultracold atomic systems \cite{daley2011quantum, smith2013quantum, cazalilla2014ultracold}, as well as developments in the control of ultracold molecular systems \cite{bohn2017cold, zeppenfeld2012sisyphus, kozyryev2017sisyphus, puri2017synthesis, wu2017cryofuge, marco2019degenerate, liu2019molecular, anderegg2019optical, chou2020frequencycomb, lin2020quantum}, a particular setting of growing interest is the spin qudit, or a multilevel quantum angular momentum degree of freedom.
+Due to advancements in experimental capabilities to address nuclear spin states (hyperfine levels) in ultracold atomic systems \cite{daley2011quantum, mischuck2012control, smith2013quantum, cazalilla2014ultracold}, as well as developments in the control of ultracold molecular systems \cite{bohn2017cold, zeppenfeld2012sisyphus, kozyryev2017sisyphus, puri2017synthesis, wu2017cryofuge, marco2019degenerate, liu2019molecular, anderegg2019optical, chou2020frequencycomb, lin2020quantum}, a particular setting of growing interest is the spin qudit, or a multilevel quantum angular momentum degree of freedom.
 Spin qudits can provide advantages over their qubit counterparts for quantum sensing \cite{hemmer2018squeezing, evrard2019enhanced}, enable quantum simulations of SU($d$) magnetism \cite{cazalilla2014ultracold, banerjee2013atomic, zhang2014spectroscopic, scazza2014observation, goban2018emergence}, and offer unique capabilities for quantum error correction \cite{albert2020robust, gross2020encoding}.

### 2020-12-01 sun_phases/spin_qudit_tomo.tex
changed/fixed old notation

@@
 
-Implicitly associating the Euler angle doublet $\v v = \p{\alpha,\beta}$ with the triplet $\p{\alpha,\beta,0}$, the transformation rules in Eq.~\eqref{eq:trans_rot_apndx} imply that we can expand the phase-space distribution of $T_{\ell m}$ as
-\begin{align}
-  \tilde T_{\ell m}\p{\v v}
+Implicitly associating the Euler angle doublet $\v v = \p{\alpha,\beta}$ with the triplet $\p{\alpha,\beta,0}$, the transformation rules in Eq.~\eqref{eq:trans_rot_apndx} imply that we can expand the phase-space representation of $T_{\ell m}$ as
+\begin{align}
+  T_{\ell m}^{\t{PS}}\p{\v v}
   \equiv \bk{s_{\v v} | T_{\ell m} | s_{\v v}}
@@
 \begin{align}
-  \tilde T_{\ell m}\p{\v v}
+  T_{\ell m}^{\t{PS}}\p{\v v}
   = \sqrt{\f{4\pi}{2s+\ell+1}

### 2020-12-03 sun_phases/spin_qudit_tomo.tex
add acknowledgments
updates and todos given feedback from DB

@@
   We consider the task of performing quantum state tomography on a $d$-state spin qudit, using only measurements of spin projection onto different quantization axes.
-  By an exact mapping onto the classical problem of signal recovery on the sphere, we prove that full reconstruction of arbitrary qudit states requires is possible with a minimal number of measurement axes, $r_d^{\t{min}}$, that is bounded by $2d-1\le r_d^{\t{min}}\le d^2$.
+  By an exact mapping onto the classical problem of signal recovery on the sphere, we prove that full reconstruction of arbitrary qudit states requires a minimal number of measurement axes, $r_d^{\t{min}}$, that is bounded by $2d-1\le r_d^{\t{min}}\le d^2$.
   We conjecture that $r_d^{\t{min}}=2d-1$, which we verify numerically for all $d\le200$.
-  We then provide algorithms with $O\p{rd^3}$ serial runtime, parallelizable down to $O\p{rd^2}$, for (i) computing (a priori) upper bounds on the expected error with which spin projection measurements along $r$ given axes can reconstruct an unknown qudit state, and (ii) estimating (a posteriori) the statistical error in a reconstructed state.
+  We then provide algorithms with $O\p{rd^3}$ serial runtime, parallelizable down to $O\p{rd^2}$, for (i) computing a priori upper bounds on the expected error with which spin projection measurements along $r$ given axes can reconstruct an unknown qudit state, and (ii) estimating a posteriori the statistical error in a reconstructed state.
   Our algorithms motivate a simple randomized tomography protocol, for which we numerically find that using more measurement axes can yield substantial benefits that plateau after $r\approx3d$.
@@
 
-Quantum state tomography, or the task of reconstructing a quantum state by collecting and processing measurement data, is an essential primitive for quantum sensing, quantum simulation, and quantum information processing.
+Quantum state tomography, the task of reconstructing a quantum state by collecting and processing measurement data, is an essential primitive for quantum sensing, quantum simulation, and quantum information processing.
 The central importance of quantum state tomography has led to the development of a variety of techniques based on compressed sensing \cite{gross2010quantum}, matrix product states \cite{cramer2010efficient}, maximum-likelihood estimation \cite{smolin2012efficient}, Bayesian inference \cite{huszar2012adaptive}, and neural networks \cite{torlai2018neuralnetwork}, among many others \cite{christandl2012reliable, qi2013quantum, opatrny1997leastsquares}.
@@
 The general problem of qudit tomography is not new, with an extensive literature on a variety of techniques \cite{manko1997spin, schmied2011tomographic, evrard2019enhanced, flammia2005minimal, thew2002qudit, salazar2012quantum, sosa-martinez2017quantum, ha2018minimal, stefano2019set, palici2020oam}.
-However, existing protocols either rely infinite-dimensional representations of a quantum spin \cite{manko1997spin, schmied2011tomographic, evrard2019enhanced}, or require the capability to perform essentially arbitrary operations on a qudit \cite{thew2002qudit, flammia2005minimal, salazar2012quantum, sosa-martinez2017quantum, ha2018minimal, stefano2019set, palici2020oam}, generally resulting in tomographic protocols that can be highly inefficient or unachievable in practice.
+However, existing protocols either rely on infinite-dimensional representations of a quantum spin \cite{manko1997spin, schmied2011tomographic, evrard2019enhanced}, or require the capability to perform essentially arbitrary operations on a qudit \cite{thew2002qudit, flammia2005minimal, salazar2012quantum, sosa-martinez2017quantum, ha2018minimal, stefano2019set, palici2020oam}, generally resulting in tomographic protocols that can be highly inefficient or unachievable in practice.
 The protocols based on infinite-dimensional representations of a quantum spin have the advantage of reconstructing its state from measurements of spin projection onto different spatial axes, which are generally accessible with any spin qudit.
-Nonetheless, these protocols obfuscate the minimal requirements for performing full state tomography, provide no straightforward error bounds or guarantees of accuracy, and (with the exception of Ref.~\cite{schmied2011tomographic}) extract only a small faction of the information contained in measurement data.
+Nonetheless, these protocols obfuscate the minimal requirements for performing full state tomography, provide no straightforward error bounds or guarantees of accuracy, and (with the exception of Ref.~\cite{schmied2011tomographic}) extract only a small fraction of the information contained in measurement data.
 
@@
 
-We begin by introducing a set of qudit operators that are analogous to the spherical harmonics $Y_{\ell m}$, and which play a central role in our work.
+% todo: define spherical harmonics
+
+We begin by introducing a set of qudit operators that are analogous to the spherical harmonic functions $Y_{\ell m}$, and which play a central role in our work.
 Consider a $d$-state spin qudit with total spin $s\equiv\frac{d-1}{2}$.
-The defining property of a spin qudit is that it describes an angular momentum degree of freedom, which has specific implications for how a spin qudit should transform under the group SO(3) of rotations in 3D space.
+The defining property of a spin qudit, distinguishing it from other qudits, is that it describes an angular momentum degree of freedom, which has specific implications for how a spin qudit should transform under the group SO(3) of rotations in 3D space.
 Due to the central importance of these transformation rules for a spin qudit, we seek a basis of operators that transform nicely under 3D rotations \footnote{Technically speaking, we seek a basis of operators that transform as an irreducible representation of SO(3).}.
@@
 where $\bk{X}_\rho\equiv\tr\p{\rho X}=\obk{\rho|X}$, and $\rho^\dag=\rho$ implies that $\rho_{\ell m}^*=\p{-1}^m\rho_{\ell,-m}$.
-The polarization operators can be physically interpreted by the transition $\ket\psi\to T_{\ell m}\ket\psi$ induced on the state $\ket\psi$ of a spin-$s$ system by the absorption of a particle with total spin $\ell$ and spin projection $m$ onto a fixed quantization axis.
+The polarization operators can be interpreted in terms of an absorption process, whereby $T_{\ell m}\ket\psi$ is (up to normalization) the state obtained after a spin-$s$ state $\ket\psi$ absorbs a particle with total spin $\ell$ and spin projection $m$ onto a fixed quantization axis.
 In analogy with the spherical harmonics $Y_{\ell m}$, we will refer to the index $\ell$ as the {\it degree} and $m$ as the {\it order} of $T_{\ell m}$.
@@
 
+% todo: add reference for Wigner function, clarify / use different language; mention Husimi distribution
+
+% todo: comment further on this representation, motivating signal recovery via the phase-space representation of qudit states
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-In fact, the problem of finding suitable axes $V$ and coefficients $C_{\ell m}\p{\v v}$ to satisfy Eq.~\eqref{eq:tomo_recovery}, solving which would enable the recovery of arbitrary spin qudit states, can be mapped onto the well-studied problem of signal recovery on the sphere \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
+In fact, the problem of finding suitable axes $V$ and coefficients $C_{\ell m}\p{\v v}$ to satisfy Eq.~\eqref{eq:tomo_recovery} can be mapped onto the well-studied problem of signal recovery on the sphere \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
 The signal recovery problem can be stated as follows: given a square-integrable function $f$ on the sphere, with the spherical harmonic expansion
@@
 Though straightforward to compute, the bound in Eq.~\eqref{eq:bound} is not tight, as it is acquired by bounding the statistical error in the empirical estimate $\tilde T_{\v v\ell,0}$ of $\bk{T_{\v v\ell,0}}_\rho$ by $\bbk{\epsilon_{\v v\ell,0}^2}\le\Gamma_\ell^2$.
-The individual bounds on $\bbk{\epsilon_{\v v\ell,0}^2}$ for each axis $\v v$ and degree $\ell$ are tight, these bounds cannot all be achieved simultaneously.
+The individual bounds on $\bbk{\epsilon_{\v v\ell,0}^2}$ for each axis $\v v$ and degree $\ell$ are tight, but these bounds cannot all be achieved simultaneously.
 There is therefore still room for improvement on the bound in Eq.~\eqref{eq:bound} by maximizing $\E_V$ over the set of all qudit states $\rho$.
@@
 The quantum error scale $\epsilon_V$ provides a pessimistic upper bound on statistical error, which can be calculated without prior knowledge of the true qudit state $\rho$.
-The actual error in the reconstruction $\tilde\rho_V$ of a particular state $\rho$ may be considerably smaller, and depends on $\rho$ itself.
+The actual shot-noise-limited error in the reconstruction $\tilde\rho_V$ of a particular state $\rho$ may be considerably smaller, and depends on $\rho$ itself.
 The exact mean squared distance between $\tilde\rho_V$ and $\rho$, derived in Appendix \ref{sec:exact}, can be written in the form
@@
 
+% todo: present protocol as an enumerated list of steps?
+
 The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and optimize (using any standard minimization algorithm) the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the quantum error scale $\epsilon_V$.
@@
 \end{align}
-The total error in $\tilde\rho_V$, measured by its root-mean-square distance from $\rho$, is approximately $\E_V\p{\tilde\rho_V}\approx\E_V\p{\rho}$.
+If reconstruction error is dominated by shot noise, then the total error in $\tilde\rho_V$, or its root-mean-square distance from $\rho$, is approximately $\E_V\p{\tilde\rho_V}\approx\E_V\p{\rho}$.
 If $\tilde\rho_V$ has negative eigenvalues, its distance from $\rho$ can be reduced with maximum-likelihood corrections \cite{smolin2012efficient}.
 
-Note that any information about $\rho$, obtained either from prior knowledge or preliminary measurement data, can be used to construct tailored or adaptive measurement protocols \cite{pereira2018adaptive} that are more efficient in terms of the number of measurements required to estimate $\rho$ to a fixed precision.
+In principle, any information about $\rho$ obtained from prior knowledge or preliminary measurement data can be used to construct tailored or adaptive measurement protocols \cite{pereira2018adaptive} that are more efficient in terms of the number of measurements required to estimate $\rho$ to a fixed precision.
 We leave the development of tailored and adaptive measurement protocols to future work.
@@
 
+% todo: include table of good measurement axes for specific values of d?
+% Yb (5/2), Cr (3), Sr (9/2), Dy (15/2?), Er (19/2)
+
 \begin{figure}[hb]
@@
   \caption{Empirical measurement-adjusted error scales $\tilde\beta(p)$ with $p$ excess measurement axes, determined by minimizing over $10^3$ choices of measurement axes $V$ or 5 minutes of runtime (for each $p$), whichever comes first.
-    Color indicates the qudit dimension $d$ in the legend.
+    Color indicates the qudit dimension $d$.
     The rapid initial drop in $\tilde\beta(p)$ implies that using more measurement axes can substantially lower the upper bound on reconstruction error provided in Eq.~\eqref{eq:bound}, and that these benefits plateau after $p\approx d$.}
@@
 % todo: write conclusions
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{acknowledgments}
+We thank Anthony M.~Polloreno and Jeremy T.~Young for helpful feedback.
+This work is supported  by the DARPA DRINQs grant, the ARO single investigator award W911NF-19-1-0210, AFOSR grant FA9550-19-1-0275, NSF grant PHY-1820885, NSF grant PHY-1734006 (JILA-PFC), and by NIST.
+\end{acknowledgments}
 
@@
 We then define the projectors $\Pi_{i\mu}\equiv\op{\mu_{\v v_i}}$ onto states $\ket{\mu_{\v v_i}}$ with definite spin projection $\mu$ along axis $\v v_i\in V$, and define $\Pi_{i\mu}^j$ to be an $N$-qudit operator with $\Pi_{i\mu}$ on the $\p{i,j}$-th tensor factor and the identity elsewhere.
-We denote the experimental outcome of measuring $\Pi_{i\mu}$ in the $\p{i,j}$-th copy of $\rho$ by $\tilde \Pi_{i\mu}^j\in\set{0,1}$.
-In other words, $\tilde \Pi_{i\mu}^j$ is the ``single-shot estimate'' of $\Pi_{i\mu}$, with $\tilde \Pi_{i\mu}^j=1$ if outcome $\mu$ was observed on the $\p{i,j}$-th experimental trial, and $\tilde \Pi_{i\mu}^j=0$ otherwise.
+We denote the experimental outcome of measuring $\Pi_{i\mu}$ in the $\p{i,j}$-th copy of $\rho$ by $\tilde\Pi_{i\mu}^j\in\set{0,1}$.
+In other words, $\tilde\Pi_{i\mu}^j$ is the ``single-shot estimate'' of $\Pi_{i\mu}$, with $\tilde\Pi_{i\mu}^j=1$ if outcome $\mu$ was observed on the $\p{i,j}$-th experimental trial, and $\tilde\Pi_{i\mu}^j=0$ otherwise.
 An empirical estimate of the expectation value $\bk{\Pi_{i\mu}}_\rho$ is provided by the fraction of times that outcome $\mu$ was observed when measuring spin projection along axis $\v v_i$, that is
 \begin{align}
-  \tilde \Pi_{i\mu} \equiv \f1n \sum_{j=1}^n \tilde \Pi_{i\mu}^j
+  \tilde\Pi_{i\mu} \equiv \f1n \sum_{j=1}^n \tilde\Pi_{i\mu}^j
   \approx \f1n \sum_{j=1}^n \tr\p{\rho^{\otimes N} \Pi_{i\mu}^j}
   = \tr\p{\rho \Pi_{i\mu}}.
-\end{align}
-For reasons that will become clear shortly, it will be useful to think of $\tilde \Pi_{i\mu}$ as an empirical estimate of $\bk{\bar \Pi_{i\mu}}_{\rho^{\otimes N}}$, where
-\begin{align}
-  \bar \Pi_{i\mu} \equiv \f1n \sum_{j=1}^n \Pi_{i\mu}^j
+  \label{eq:proj_estimate}
+\end{align}
+For reasons that will be clarified shortly, it will be useful to think of $\tilde\Pi_{i\mu}$ as an empirical estimate of $\bk{\bar\Pi_{i\mu}}_{\rho^{\otimes N}}$, where
+\begin{align}
+  \bar\Pi_{i\mu} \equiv \f1n \sum_{j=1}^n \Pi_{i\mu}^j
   \label{eq:mean_proj}
@@
 is the average of $\Pi_{i\mu}$ applied to all copies of $\rho$ for which spin projection is measured along the axis $\v v_i$.
+Eq.~\eqref{eq:proj_estimate} implies that
+\begin{align}
+  \tilde\Pi_{i\mu}
+  \approx \bk{\bar\Pi_{i\mu}}_{\rho^{\otimes N}}
+  = \bk{\Pi_{i\mu}}_\rho.
+\end{align}
 
@@
   \epsilon_{i\mu}
-  \equiv \tilde \Pi_{i\mu} - \bk{\Pi_{i\mu}}_{\rho}
-  = \tilde \Pi_{i\mu} - \bk{\bar\Pi_{i\mu}}_{\rho^{\otimes N}}
+  \equiv \tilde\Pi_{i\mu} - \bk{\Pi_{i\mu}}_{\rho}
+  = \tilde\Pi_{i\mu} - \bk{\bar\Pi_{i\mu}}_{\rho^{\otimes N}}
 \end{align}
@@
 where we used the fact that the covariance $\cov_\rho\p{X,Y}$ is linear in both $X$ and $Y$.
-Due to the appearance of $\delta_{ii'}$ above and the orthogonality of rotated polarization operators $T_{i\ell}$ and $T_{i'\ell'}$ with degrees $\ell\ne\ell'$, only the variance $\bbk{\epsilon_{i\ell}^2}$ ultimately contributes to reconstruction error.
+Due to the appearance of $\delta_{ii'}$ above and the orthogonality of polarization operators $T_{i\ell}$ and $T_{i'\ell'}$ with degrees $\ell\ne\ell'$, it turns out that only the variances $\bbk{\epsilon_{i\ell}^2}$ will ultimately contribute to reconstruction error (see Appendix \ref{sec:revisiting}).
 We therefore seek to find an upper bound on $\bbk{\epsilon_{i\ell}^2}$.
@@
   &&
-  \bbk{\epsilon_{i\ell}} \le \f1{2n}.
+  \bbk{\epsilon_{i\ell}^2} \le \f1{2n}.
 \end{align}
@@
   f_{\ell_1 m_1;\ell_2 m_2}^{LM}
-  \equiv \obk{T_{LM} | T_{\ell_1 m_1} T_{\ell_2 m_2}}.
-\end{align}
+  \equiv \obk{T_{LM} | T_{\ell_1 m_1} T_{\ell_2 m_2}},
+\end{align}
+which allow us to simplify the commutators in Eq.~\eqref{eq:spin_trans} of Appendix \ref{sec:rotations}.
 Using the symmetry properties of Clebsch-Gordan coefficients, namely

### 2020-12-04 sun_phases/spin_qudit_tomo.tex
some changes from review/feedback

@@
 
-% todo: define spherical harmonics
-
-We begin by introducing a set of qudit operators that are analogous to the spherical harmonic functions $Y_{\ell m}$, and which play a central role in our work.
+We begin by introducing a set of qudit operators that are a quantum analogue of the spherical harmonics (in a sense that will be clarified below), and which play a central role in our work.
 Consider a $d$-state spin qudit with total spin $s\equiv\frac{d-1}{2}$.
@@
 where $\bk{s\mu;\ell m|s\nu}$ is a Clebsh-Gordan coefficient that enforces $\ell\in\set{0,1,\cdots,d-1}$ and $m\in\set{-\ell,-\ell+1,\cdots,\ell}$, such that there are $d^2$ polarization operators in total.
-For brevity, we will generally treat the value of $d$ (or $s$) as constant but arbitrary throughout this work, and suppress any explicit dependence of quantities or operators such as $T_{\ell m}$ on $d$.
+For brevity, we will generally treat the value of $d$ as constant but arbitrary throughout this work, and we will suppress any explicit dependence of quantities or operators such as $T_{\ell m}$ on $d$.
 The polarization operators are orthonormal with respect to the trace inner product, and transform nicely under conjugation:
@@
 The polarization operators can be interpreted in terms of an absorption process, whereby $T_{\ell m}\ket\psi$ is (up to normalization) the state obtained after a spin-$s$ state $\ket\psi$ absorbs a particle with total spin $\ell$ and spin projection $m$ onto a fixed quantization axis.
-In analogy with the spherical harmonics $Y_{\ell m}$, we will refer to the index $\ell$ as the {\it degree} and $m$ as the {\it order} of $T_{\ell m}$.
+In analogy with the complex spherical harmonics $Y_{\ell m}$, we will refer to $\ell$ as the {\it degree} and $m$ as the {\it order} of $T_{\ell m}$.
 
 The polarization operators are spherical tensor operators, whose degree is preserved under 3D rotations generated by the spin operators $S_\x,S_\y,S_\z$.
-Moreover, the degree-$\ell$ polarization operators $T_{\ell m}$ transform identically to spin-$\ell$ particles under 3D rotations (see Appendix \ref{sec:rotations}).
+Moreover, the degree-$\ell$ polarization operators $T_{\ell m}$ transform identically to spin-$\ell$ particles and spherical harmonics $Y_{\ell m}$ under 3D rotations (see Appendix \ref{sec:rotations}).
 For any triplet of Euler angles $\v\omega=\p{\alpha,\beta,\gamma}$, we can therefore define the rotation operator
@@
 
-As a final point, we mention that the polarization operators $T_{\ell m}$ share a connection to the spherical harmonics $Y_{\ell m}$ that goes beyond the rules for their transformation under 3D rotations.
-In fact, the phase-space (Wigner function) representation of $T_{\ell m}$ is proportional to $Y_{\ell m}$.
-The phase-space representation of a spin qudit operator $X$ assigns, to each point $\v v$ on the sphere, the number
+The polarization operators $T_{\ell m}$ share a connection to the spherical harmonics $Y_{\ell m}$ that goes beyond their indexing and the rules for their transformation under 3D rotations.
+In fact, the phase-space representation of $T_{\ell m}$ is proportional to $Y_{\ell m}$.
+The phase-space representation of a spin qudit operator $X$ assigns, to each point $\v v$ on the sphere, the complex number
 \begin{align}
@@
 \end{align}
-where the scalar $c_\ell$ simply enforces that $\obk{T_{\ell m}|T_{\ell m}}=1$.
-The polarization operators $T_{\ell m}$ are thus a quantum analogue of spherical harmonics $Y_{\ell m}$, which play an important role in phase-space formalisms for spin qudits \cite{li2013weylwignermoyal}.
-
-% todo: add reference for Wigner function, clarify / use different language; mention Husimi distribution
-
-% todo: comment further on this representation, motivating signal recovery via the phase-space representation of qudit states
+where the scalar $c_\ell$ simply enforces $\obk{T_{\ell m}|T_{\ell m}}=1$.
+The polarization operators $T_{\ell m}$ are thus a quantum analogue of the spherical harmonics $Y_{\ell m}$, and play an important role in phase-space formalisms for spin qudits \cite{li2013weylwignermoyal}.
+
+As a special case, the phase-space representation $\rho^{\t{PS}}$ of a qudit state $\rho$ is common known as its Husimi distribution.
+Performing tomography on an unknown qudit state $\rho$ is therefore equivalent to reconstructing the unknown distribution $\rho^{\t{PS}}$ on the sphere.
+In principle, the representation $\rho^{\t{PS}}$ of a finite-dimensional qudit state $\rho$ can be reconstructed from the values $\rho^{\t{PS}}\p{\v v}=\bk{s_{\v v}|\rho|s_{\v v}}$ at a finite number of points $\v v$ on the sphere.
+In practice, the value $\bk{s_{\v v}|\rho|s_{\v v}}$ is determined by measuring spin projection along $\v v$, which also provides measurement data on all spin projections $\bk{\mu_{\v v}|\rho|\mu_{\v v}}$ with $\mu\in\set{s,s-1,\cdots,-s}$ and $\ket{\mu_{\v v}}\equiv R\p{\v v}\ket{\mu}$; one would like to make use of this additional data as well.
+We clarify the connection between the quantum problem of reconstructing $\rho$ from spin projection measurements and the classical problem of reconstructing $\rho^{\t{PS}}$ from its values $\rho^{\t{PS}}\p{\v v}$ in the following section.
+In addition to providing a new perspective on the problem of spin qudit tomography, the connection will allow us to place upper bounds on the minimal number of measurement axes $\v v$ necessary for full recovery of an unknown state.
 
@@
   T_{\ell m}
-  = \sum_{\v v,k,n} C_{\ell mk}\p{\v v}
-  D^k_{n,0}\p{\v v} T_{kn}.
-\end{align}
-Orthogonality of the polarization operators motivates the ansatz $C_{\ell mk}\p{\v v}=\delta_{k\ell}C_{\ell m}\p{\v v}$, which in turn implies that
-\begin{align}
-  \sum_{\v v} C_{\ell m}\p{\v v} D^\ell_{n,0}\p{\v v} = \delta_{mn}
+  = \sum_{\v v,k,n} C_{\ell mk}\p{\v v} D^k_{n,0}\p{\v v} T_{kn}.
+\end{align}
+Orthogonality of the polarization operators then implies the decomposition $C_{\ell mk}\p{\v v}=\delta_{\ell k}C_{\ell m}\p{\v v}$, and in turn
+\begin{align}
+  \sum_{\v v} C_{\ell m}\p{\v v} D^\ell_{n,0}\p{\v v}
+  = \delta_{mn}
   \label{eq:tomo_recovery}
@@
 for all $\ell$.
-
 In fact, the problem of finding suitable axes $V$ and coefficients $C_{\ell m}\p{\v v}$ to satisfy Eq.~\eqref{eq:tomo_recovery} can be mapped onto the well-studied problem of signal recovery on the sphere \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
@@
   \sum_{\v v} \tilde C_{\ell m}\p{\v v} Y_{kn}\p{\v v}
-  = \delta_{k\ell} \delta_{mn},
+  = \delta_{\ell k} \delta_{mn},
   \label{eq:full_recovery}
@@
 \item there are at most $2d-1$ polarization operators of any given degree, and
-\item measuring spin projection along any single axis provides measurement data for polarization operators of all degrees,
+\item measuring spin projection along a single axis provides measurement data for polarization operators of all degrees,
 \end{enumerate*}
 one might expect that $2d-1$ suitably chosen measurement axes would always suffice to perform full tomography of a $d$-level spin qudit.
-In the language of classical signal recovery, we would expect there to exist a set of $2\ell_{\t{max}}+1$ points on the sphere that enable the recovery of any function with fixed degree $\ell\le\ell_{\t{max}}$, given prior knowledge of $\ell$ \footnote{We could further conjecture that if {\it any} set of points $V\subset\Omega_2$ enables the recovery of degree-$\ell_{\t{max}}$ functions on the sphere, then $V$ enables the recovery degree-$\ell$ functions with $\ell\le\ell_{\t{max}}$.
+In the language of classical signal recovery, we would expect there to exist a set of $2\ell_{\t{max}}+1$ points on the sphere that enable the recovery of any function with fixed degree $\ell\le\ell_{\t{max}}$, given prior knowledge of $\ell$ \footnote{We could further conjecture that if {\it any} set of points $V$ enables the recovery of degree-$\ell_{\t{max}}$ functions on the sphere, then $V$ enables the recovery degree-$\ell$ functions with $\ell\le\ell_{\t{max}}$.
   However, this condition is stronger than necessary to guarantee full recovery of a $d$-level spin qudit state with $2d-1$ measurement axes, which only requires the existence of {\it some} such set of points.}.
 Unfortunately, we have no general proof of this conjecture, which requires a deeper analysis of fixed-degree signal recovery on the sphere.
-Nonetheless, in the following sections we numerically verify that $2d-1$ points suffice for all $d\le200$.
+Nonetheless, in the following section we numerically verify that $2d-1$ points suffice for all $d\le200$.
 
@@
 
-The reconstruction error bound in Eq.~\eqref{eq:bound} nominally provides a straightforward answer: at a fixed total number of measurements, $N=n\abs{V}$, the number of measurement axes should be chosen to minimize the (squared) reconstruction error $\E_V\p{\rho}^2<\epsilon_V^2/n\propto\epsilon_V^2\abs{V}$.
-We therefore consider the measurement-adjusted error scale
-\begin{align}
-  \beta(p)
-  \equiv \min\set{\epsilon_V:\abs{V}=2d-1+p} \times \sqrt{2d-1+p}
-\end{align}
-as a function of $p$, the number of measurement axes exceeding $2d-1$.
-Though we cannot minimize over all suitable choices of measurement axes $V$ to compute $\beta(p)$, we can compute an empirical upper bound $\tilde\beta(p)\ge\beta(p)$ by minimizing over a large number of randomly chosen $V$.
-Figure \ref{fig:axes} shows the results of such empirical minimization, where we find that $\tilde\beta(p)$ drops substantially before plateauing at $p\approx d$, after which there are only minor benefits to using more measurement axes.
-In the interest of reducing experimental complexity as well as the runtime of our tomography protocol, which grow linearly in $\abs{V}$, we therefore conclude that spin qudit tomography should be performed with $\abs{V}\approx3d$ measurement axes.
-
-% todo: include table of good measurement axes for specific values of d?
-% Yb (5/2), Cr (3), Sr (9/2), Dy (15/2?), Er (19/2)
-
-\begin{figure}[hb]
+\begin{figure}
   \centering
@@
 
+The reconstruction error bound in Eq.~\eqref{eq:bound} nominally provides a straightforward answer: at a fixed total number of measurements, $N=n\abs{V}$, the number of measurement axes should be chosen to minimize the (squared) reconstruction error $\E_V\p{\rho}^2<\epsilon_V^2/n\propto\epsilon_V^2\abs{V}$.
+We therefore consider the measurement-adjusted error scale
+\begin{align}
+  \beta(p)
+  \equiv \min\set{\epsilon_V:\abs{V}=2d-1+p} \times \sqrt{2d-1+p}
+\end{align}
+as a function of $p$, the number of measurement axes exceeding $2d-1$.
+Though we cannot minimize over all suitable choices of measurement axes $V$ to compute $\beta(p)$, we can compute an empirical upper bound $\tilde\beta(p)\ge\beta(p)$ by minimizing over a large number of randomly chosen $V$.
+Figure \ref{fig:axes} shows the results of such empirical minimization, where we find that $\tilde\beta(p)$ drops substantially before plateauing at $p\approx d$, after which there are only minor benefits to using more measurement axes.
+In the interest of reducing experimental complexity as well as the runtime of our tomography protocol, which grow linearly in $\abs{V}$, we therefore conclude that spin qudit tomography should be performed with $\abs{V}\approx3d$ measurement axes.
+
+% todo: include table of good measurement axes for specific values of d?
+% Yb (5/2), Cr (3), Sr (9/2), Dy (15/2?), Er (19/2)
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \end{align}
-In this way, the polarization operators are a quantum analogue of spherical harmonics.
+In this way, the polarization operators are a quantum analogue of the spherical harmonics.
 

### 2020-12-04 sun_phases/spin_qudit_tomo.tex
wrote tomography protocol as an enumerated list of steps minor changes to appendices

@@
 
-% todo: present protocol as an enumerated list of steps?
-
-The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and optimize (using any standard minimization algorithm) the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the quantum error scale $\epsilon_V$.
-If $\abs{V}$ is too large for such optimization to be practical, one can instead generate many random sets of measurement axes, and then choose the set with the smallest quantum error scale $\epsilon_V$.
-After choosing a set of measurement axes $V$, make $n$ measurements of spin projection along every axis $\v v\in V$, and set $\tilde\Pi_{\v v\mu}\approx\bk{\Pi_{\v v\mu}}_\rho$ to the fraction of times in which the measurement outcome was $\mu$.
-The estimates $\tilde\Pi_{\v v\mu}$ of $\bk{\Pi_{\v v\mu}}_\rho$ determine estimates
+The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following protocol for spin qudit tomography:
+\begin{enumerate}
+\item Select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and use any standard minimization algorithm to optimize the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) by minimizing the quantum error scale $\epsilon_V$.
+  If $\abs{V}$ is too large for such optimization to be practical, you can simply generate many random sets of measurement axes, and then choose the set with the smallest quantum error scale $\epsilon_V$.
+\item For each axis $\v v\in V$, make $n$ measurements of spin projection, and set $\tilde\Pi_{\v v\mu}\approx\bk{\Pi_{\v v\mu}}_\rho$ to the fraction of times in which the measurement outcome was $\mu$.
+\item Use the the estimates $\tilde\Pi_{\v v\mu}$ of $\bk{\Pi_{\v v\mu}}_\rho$ to compute the estimates
 \begin{align}
@@
 \end{align}
-of $\bk{T_{\v v\ell,0}}_\rho$, which can in turn be used to construct the estimate $\tilde\rho_V$ of $\rho$.
-Specifically, the $2\ell+1$ (normalized) right singular vectors $\v x^V_{\ell k} = \sum_{\v v} x^V_{\ell k \v v} \ket{\v v}$ and corresponding singular values $M^V_{\ell k}$ of the matrices $M_{V\ell}$ defined in Eq.~\eqref{eq:block} determine the components
-\begin{align}
-  \tilde\rho^V_{\ell k}
-  = \f1{M^V_{\ell k}} \sum_{\v v}
+of $\bk{T_{\v v\ell,0}}_\rho$, where the matrix elements of $T_{\ell,0}$ are provided in Eq.~\eqref{eq:trans_op}.
+\item For each $\ell\in\set{0,1,\cdots,d-1}$, build the measurement matrix $M_{V\ell}$ defined in Eq.~\eqref{eq:block}, and compute its singular value decomposition to identify the singular values $M^V_{\ell k}$ and corresponding right singular vectors $\v x^V_{\ell k} = \sum_{\v v} x^V_{\ell k \v v} \ket{\v v}$, indexed by an integer $k\in\set{1,2,\cdots,2\ell+1}$.
+% todo: cite codes
+\item Compute the complex numbers
+  \begin{align}
+    \tilde\rho^V_{\ell k} \equiv \f1{M^V_{\ell k}} \sum_{\v v}
     x^V_{\ell k\v v} \, \tilde T_{\v v\ell,0},
   \end{align}
-of
+  and the operators
+  \begin{align}
+    Q_{\ell k} \equiv \f1{M^V_{\ell k}} \sum_{\v v}
+    x^V_{\ell k\v v} T_{\v v\ell,0},
+  \end{align}
+  and combine them into the estimate
   \begin{align}
     \tilde\rho_V = \sum_{\ell,k} \tilde\rho^V_{\ell k} Q_{\ell k}
-  \approx \rho,
-  &&
-  Q_{\ell k} = \f1{M^V_{\ell k}} \sum_{\v v}
-  x^V_{\ell k\v v} T_{\v v\ell,0}.
-\end{align}
+    \approx \rho.
+  \end{align}
+\end{enumerate}
 If reconstruction error is dominated by shot noise, then the total error in $\tilde\rho_V$, or its root-mean-square distance from $\rho$, is approximately $\E_V\p{\tilde\rho_V}\approx\E_V\p{\rho}$.
@@
 
-The tomography protocol we outline above leaves open the question of {\it how many} measurement axes to use.
+The tomography protocol outlined above leaves open the question of {\it how many} measurement axes to use.
 Though $2d-1$ measurement axes may be sufficient to perform full state tomography, this is not necessarily the best choice of $\abs{V}$.
@@
 Though we cannot minimize over all suitable choices of measurement axes $V$ to compute $\beta(p)$, we can compute an empirical upper bound $\tilde\beta(p)\ge\beta(p)$ by minimizing over a large number of randomly chosen $V$.
-Figure \ref{fig:axes} shows the results of such empirical minimization, where we find that $\tilde\beta(p)$ drops substantially before plateauing at $p\approx d$, after which there are only minor benefits to using more measurement axes.
+Figure \ref{fig:axes} shows the results of such empirical minimization, where we find that $\tilde\beta(p)$ drops substantially with $p$ before plateauing at $p\approx d$, after which there are only minor benefits to using more measurement axes.
 In the interest of reducing experimental complexity as well as the runtime of our tomography protocol, which grow linearly in $\abs{V}$, we therefore conclude that spin qudit tomography should be performed with $\abs{V}\approx3d$ measurement axes.
 
-% todo: include table of good measurement axes for specific values of d?
+% todo: include table of good measurement axes for specific values of d
 % Yb (5/2), Cr (3), Sr (9/2), Dy (15/2?), Er (19/2)
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \section{Conclusions}
-
-% todo: write conclusions
 
@@
 \begin{align}
-  \rho_i^\star = \f12\p{\Pi_{i\mu_{\t{max}}} + \Pi_{i\mu_{\t{min}}}}.
-\end{align}
-To find a nicer analytical bound on $\bbk{\epsilon_{i\ell}^2}$, we can use normalization of the polarization operators, $\obk{T_{i\ell}|T_{i\ell}}=\sum_\mu t_{\ell\mu}^2 = 1$, and the fact that all probabilities $p_\mu\le1$ to bound
+  \rho_i^\star = \f12\p{\Pi_{i\mu_{\t{max}}} + \Pi_{i\mu_{\t{min}}}},
+\end{align}
+where $\mu_{\t{max}}$ ($\mu_{\t{min}}$) is the index that maximizes (minimizes) $t_{\ell\mu}$.
+
+To find an analytical bound on $\bbk{\epsilon_{i\ell}^2}$ that is easier to interpret, we can use normalization of the polarization operators, $\obk{T_{i\ell}|T_{i\ell}}=\sum_\mu t_{\ell\mu}^2 = 1$, and the fact that all probabilities $p_\mu\le1$ to bound
 \begin{align}
@@
 \end{align}
-We can get a tighter analytical bound on $\bbk{\epsilon_{i\ell}^2}$ by considering the fact that $t_{\ell\mu}^2 = t_{\ell,-\mu}^2$ due to the symmetries of the Clebsch-Gordan coefficients.
-Denoting the index that maximizes $t_{\ell\mu}^2$ by $\mu_{\t{max}}$, it follows that if $\mu_{\t{max}}\ne0$ then
+We can get a tighter bound by considering the fact that $t_{\ell\mu}^2 = t_{\ell,-\mu}^2$ due to the symmetries of the Clebsch-Gordan coefficients.
+It follows that if $\mu_{\t{max}}\ne0$ then
 \begin{align}
@@
 \end{align}
-where $\mu_{\t{min}}$ is the index that minimizes $t_{\ell\mu}^2$.
-This inequality lets us bound
+which lets us bound
 \begin{align}
@@
 \end{align}
-If is straightforward to show that $\lambda\p{x}$ is maximally $\lambda^\star\equiv \max_x \lambda\p{x} = \sqrt{3/8}$, so
+It is straightforward to show that $\lambda\p{x}$ is maximally $\lambda^\star\equiv \max_x \lambda\p{x} = \sqrt{3/8}$, so
 \begin{align}
@@
 In principle, maximizing the reconstruction error in Eq.~\eqref{eq:error_SM} over all qudit states $\rho$ would provide a tight upper bound on reconstruction error for any set of axes $V$.
-To simplify this task somewhat, we first maximize Eq.~\eqref{eq:error_SM} over all $\rho$ with $\tr\p\rho=1$: this maximum occurs at a ``state'' $\sigma_V^\star$ whose components are defined by
+To simplify this task somewhat, we first maximize Eq.~\eqref{eq:error_SM} over all $\rho$ with $\tr\p\rho=1$: this maximum occurs at a ``state'' $\sigma_V^\star$ whose components are given by
 \begin{align}

### 2020-12-04 sun_phases/spin_qudit_tomo.tex
removed redundant step from tomography protocol

@@
 
-As a special case, the phase-space representation $\rho^{\t{PS}}$ of a qudit state $\rho$ is common known as its Husimi distribution.
+As a special case, the phase-space representation $\rho^{\t{PS}}$ of a qudit state $\rho$ is commonly known as its Husimi distribution.
 Performing tomography on an unknown qudit state $\rho$ is therefore equivalent to reconstructing the unknown distribution $\rho^{\t{PS}}$ on the sphere.
@@
 \begin{enumerate}
-\item Select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and use any standard minimization algorithm to optimize the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) by minimizing the quantum error scale $\epsilon_V$.
-  If $\abs{V}$ is too large for such optimization to be practical, you can simply generate many random sets of measurement axes, and then choose the set with the smallest quantum error scale $\epsilon_V$.
+\item Select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and use any standard minimization algorithm to optimize the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) by minimizing the quantum error scale $\epsilon_V$ in Eq.~\eqref{eq:bound}.
+  If $\abs{V}$ is too large for such optimization, you can simply generate many sets of random measurement axes, and then choose the set with the smallest quantum error scale $\epsilon_V$.
+  Note that computing the error scale $\epsilon_V$ requires, for each $\ell\in\set{0,1,\cdots,d-1}$, constructing the measurement matrix $M_{V\ell}$ in Eq.~\eqref{eq:block} and computing its singular value decomposition.
+  Save all measurement matrix data associated with the final measurement axes $V$ for later use. % todo: cite codes
 \item For each axis $\v v\in V$, make $n$ measurements of spin projection, and set $\tilde\Pi_{\v v\mu}\approx\bk{\Pi_{\v v\mu}}_\rho$ to the fraction of times in which the measurement outcome was $\mu$.
@@
 of $\bk{T_{\v v\ell,0}}_\rho$, where the matrix elements of $T_{\ell,0}$ are provided in Eq.~\eqref{eq:trans_op}.
-\item For each $\ell\in\set{0,1,\cdots,d-1}$, build the measurement matrix $M_{V\ell}$ defined in Eq.~\eqref{eq:block}, and compute its singular value decomposition to identify the singular values $M^V_{\ell k}$ and corresponding right singular vectors $\v x^V_{\ell k} = \sum_{\v v} x^V_{\ell k \v v} \ket{\v v}$, indexed by an integer $k\in\set{1,2,\cdots,2\ell+1}$.
-% todo: cite codes
-\item Compute the complex numbers
-  \begin{align}
-    \tilde\rho^V_{\ell k} \equiv \f1{M^V_{\ell k}} \sum_{\v v}
-    x^V_{\ell k\v v} \, \tilde T_{\v v\ell,0},
-  \end{align}
-  and the operators
-  \begin{align}
-    Q_{\ell k} \equiv \f1{M^V_{\ell k}} \sum_{\v v}
+\item Denoting the nonzero singular values of $M_{V\ell}$ by $M^V_{\ell k}$ and the corresponding right singular vectors by $\v x^V_{\ell k} = \sum_{\v v} x^V_{\ell k \v v} \ket{\v v}$, compute the coefficients and operators
+  \begin{align}
+    \tilde\rho^V_{\ell k} &\equiv \f1{M^V_{\ell k}} \sum_{\v v}
+    x^V_{\ell k\v v} \, \tilde T_{\v v\ell,0}, \\
+    Q_{\ell k} &\equiv \f1{M^V_{\ell k}} \sum_{\v v}
     x^V_{\ell k\v v} T_{\v v\ell,0},
@@
 \end{enumerate}
-If reconstruction error is dominated by shot noise, then the total error in $\tilde\rho_V$, or its root-mean-square distance from $\rho$, is approximately $\E_V\p{\tilde\rho_V}\approx\E_V\p{\rho}$.
+If reconstruction error is dominated by shot noise, then the total error in $\tilde\rho_V$, or its root-mean-square distance from $\rho$, is approximately $\E_V\p{\tilde\rho_V}\approx\E_V\p{\rho}$, which can be computed from Eq.~\eqref{eq:error}.
+% todo: cite codes
 If $\tilde\rho_V$ has negative eigenvalues, its distance from $\rho$ can be reduced with maximum-likelihood corrections \cite{smolin2012efficient}.

### 2020-12-04 sun_phases/spin_qudit_tomo.tex
minor change to wording in paragraph about Husimi distribution

@@
 Performing tomography on an unknown qudit state $\rho$ is therefore equivalent to reconstructing the unknown distribution $\rho^{\t{PS}}$ on the sphere.
-In principle, the representation $\rho^{\t{PS}}$ of a finite-dimensional qudit state $\rho$ can be reconstructed from the values $\rho^{\t{PS}}\p{\v v}=\bk{s_{\v v}|\rho|s_{\v v}}$ at a finite number of points $\v v$ on the sphere.
+In principle, the representation $\rho^{\t{PS}}$ of a finite-dimensional qudit state $\rho$ can be reconstructed from the values $\rho^{\t{PS}}\p{\v v}=\bk{s_{\v v}|\rho|s_{\v v}}$ at a finite number of points $\v v$.
 In practice, the value $\bk{s_{\v v}|\rho|s_{\v v}}$ is determined by measuring spin projection along $\v v$, which also provides measurement data on all spin projections $\bk{\mu_{\v v}|\rho|\mu_{\v v}}$ with $\mu\in\set{s,s-1,\cdots,-s}$ and $\ket{\mu_{\v v}}\equiv R\p{\v v}\ket{\mu}$; one would like to make use of this additional data as well.
 We clarify the connection between the quantum problem of reconstructing $\rho$ from spin projection measurements and the classical problem of reconstructing $\rho^{\t{PS}}$ from its values $\rho^{\t{PS}}\p{\v v}$ in the following section.
-In addition to providing a new perspective on the problem of spin qudit tomography, the connection will allow us to place upper bounds on the minimal number of measurement axes $\v v$ necessary for full recovery of an unknown state.
+In addition to providing a new perspective on the problem of spin qudit tomography, this connection will allow us to place upper bounds on the minimal number of measurement axes $\v v$ necessary for full recovery of an unknown state.
 

### 2020-12-04 sun_phases/spin_qudit_tomo.tex
fixed variety of errors involving rotation matrices vs. their conjugates

@@
 The polarization operators are spherical tensor operators, whose degree is preserved under 3D rotations generated by the spin operators $S_\x,S_\y,S_\z$.
-Moreover, the degree-$\ell$ polarization operators $T_{\ell m}$ transform identically to spin-$\ell$ particles and spherical harmonics $Y_{\ell m}$ under 3D rotations (see Appendix \ref{sec:rotations}).
-For any triplet of Euler angles $\v\omega=\p{\alpha,\beta,\gamma}$, we can therefore define the rotation operator
+Moreover, the degree-$\ell$ polarization operators $T_{\ell m}$ transform similarly to spin-$\ell$ particles and spherical harmonics $Y_{\ell m}$ under 3D rotations (see Appendix \ref{sec:rotations}).
+Specifically, for any triplet of angles $\v\omega=\p{\alpha,\beta,\gamma}$, we can therefore define the rotation operator
 \begin{align}
@@
 \end{align}
-and expand polarization operators in a rotated frame as
+and expand rotated polarization operators as
 \begin{align}
   T_{\v\omega\ell m} \equiv
-  R\p{\v\omega}^\dag T_{\ell m} R\p{\v\omega}
-  = \sum_{n=-\ell}^\ell D_{nm}^\ell\p{\v\omega} T_{\ell n},
+  R\p{\v\omega} T_{\ell m} R\p{\v\omega}^\dag
+  = \sum_{n=-\ell}^\ell D_{mn}^\ell\p{\bar{\v\omega}}^* T_{\ell n},
   \label{eq:trans_rot}
 \end{align}
-where
-\begin{align}
-  D_{mn}^\ell\p{\v\omega} \equiv \bk{\ell m|R\p{\v\omega}|\ell n}
+where $\bar{\v\omega}=\p{\gamma,\beta,\alpha}$ is the reversal of $\v\omega$, and
+\begin{align}
+  D_{mn}^\ell\p{\bar{\v\omega}}
+  \equiv \bk{\ell m|R\p{\bar{\v\omega}}|\ell n}
 \end{align}
@@
 Throughout this work, we will primarily consider rotations of the sphere that take the north pole to a point $\v v=\p{\alpha,\beta}$ at azimuthal angle $\alpha$ and polar angle $\beta$.
-For ease of notation, we will therefore implicitly associate Euler angle doublets $\v v=\p{\alpha,\beta}$ with the triplet $\p{\alpha,\beta,0}$, such that $R\p{\v v} \equiv R\p{\alpha,\beta,0}$, $D^\ell_{mn}\p{\v v} \equiv D^\ell_{mn}\p{\alpha,\beta,0}$, and $T_{\v v\ell m} \equiv T_{\p{\alpha,\beta,0},\ell m}$.
-
-The polarization operators $T_{\ell m}$ share a connection to the spherical harmonics $Y_{\ell m}$ that goes beyond their indexing and the rules for their transformation under 3D rotations.
+For ease of notation, we therefore define $R\p{\v v} \equiv R\p{\alpha,\beta,0}$, $T_{\v v\ell m} \equiv T_{\p{\alpha,\beta,0},\ell m}$, and $D^\ell_{mn}\p{\v v} \equiv D^\ell_{mn}\p{0,\beta,\alpha}$.
+
+The polarization operators $T_{\ell m}$ share a connection to the spherical harmonics $Y_{\ell m}$ that goes beyond the rules for their transformation under 3D rotations.
 In fact, the phase-space representation of $T_{\ell m}$ is proportional to $Y_{\ell m}$.
@@
   T_{\ell m}
-  = \sum_{\v v,k,n} C_{\ell mk}\p{\v v} D^k_{n,0}\p{\v v} T_{kn}.
+  = \sum_{\v v,k,n} C_{\ell mk}\p{\v v} D^k_{0,n}\p{\v v}^* T_{kn}.
 \end{align}
@@
 \begin{align}
-  \sum_{\v v} C_{\ell m}\p{\v v} D^\ell_{n,0}\p{\v v}
+  \sum_{\v v} C_{\ell m}\p{\v v} D^\ell_{0,n}\p{\v v}^*
   = \delta_{mn}
@@
 for all $\ell$.
+
 In fact, the problem of finding suitable axes $V$ and coefficients $C_{\ell m}\p{\v v}$ to satisfy Eq.~\eqref{eq:tomo_recovery} can be mapped onto the well-studied problem of signal recovery on the sphere \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
@@
 We will refer to Eq.~\eqref{eq:full_recovery} as the {\it full recovery problem}, and Eq.~\eqref{eq:tomo_recovery} as the the {\it reduced recovery problem}.
-Due to the fact that $D^k_{n,0}\p{\v v} = \sqrt{\frac{4\pi}{2\ell+1}}\, Y_{kn}\p{\v v}^*$, any solution to the full recovery problem automatically solves the reduced recovery problem by setting $C_{\ell m}\p{\v v} = \sqrt{\frac{2\ell+1}{4\pi}}\, \tilde C_{\ell m}\p{\v v}^*$.
+Due to the fact that $D^\ell_{0,m}\p{\v v} = \sqrt{\frac{4\pi}{2\ell+1}}\, Y_{\ell m}\p{\v v}$, any solution to the full recovery problem automatically solves the reduced recovery problem by setting $C_{\ell m}\p{\v v} = \sqrt{\frac{2\ell+1}{4\pi}}\, \tilde C_{\ell m}\p{\v v}^*$.
 In principle, this mapping allows us to import a host of existing signal recovery algorithms \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality} for the task of spin qudit tomography.
@@
 If the function $f$ is {\it band-limited} at degree $L$, which is to say that $f_{\ell m}=0$ for all $\ell\ge L$, then the full recovery problem in Eq.~\eqref{eq:full_recovery} is provably solvable with a suitable choice of $\abs{V}=L^2$ points on the sphere \cite{freeden2008spherical, freeden2018spherical}.
-The existence of these solutions to the full recovery problem in turn implies the existence of $d^2$ measurement axes that suffice to reconstruct arbitrary states of $d$-level spin qudit, whose possible states are essentially band-limited at degree $d$ \footnote{Specifically, the phase-space representation of a $d$-level spin qudit state $\rho$ is band-limited at degree $d$.}.
+The existence of these solutions to the full recovery problem in turn implies the existence of $d^2$ measurement axes that suffice to reconstruct arbitrary states of $d$-level spin qudit, whose possible states (or rather, their phase-space representations) are band-limited at degree $d$.
 Moreover, for any fixed degree $\ell$, finding solutions to the reduced recovery problem in Eq.~\eqref{eq:tomo_recovery} is equivalent to the recovery of a degree-$\ell$ function $f_\ell = \sum_m f_{\ell m} Y_{\ell m}$, which is provably possible with $\abs{V}=2\ell+1$ samples \cite{freeden2008spherical}.
@@
 \begin{align}
-  M_V \equiv \sum_{\v v,\ell} \oket{T_{\v v\ell,0}} \bra{\v v\ell}.
+  M_V \equiv \sum_{\v v,\ell} \ket{\v v\ell} \obra{T_{\v v\ell,0}}.
   \label{eq:meas_mat}
 \end{align}
-Here $\v v$ and $\ell$ label a column of $M_V$, or equivalently label a standard basis vector $\ket{\v v\ell}$ of a $\p{\abs{V}\times d}$-dimensional vector space.
+Here $\v v$ and $\ell$ label a row of $M_V$, or equivalently label a standard basis vector $\ket{\v v\ell}$ of a $\p{\abs{V}\times d}$-dimensional vector space, and $\obra{T_{\v v\ell,0}}$ is the conjugate transpose of $\oket{T_{\v v\ell,0}}$.
 A necessary and sufficient condition for $V$ to allow for full state tomography is that the measured polarization operators in $\set{T_{\v v\ell,0}}$, or equivalently the columns of $M_V$, span the entire ($d^2$-dimensional) space of operators on a $d$-level spin qudit.
 In this case $M_V$ must be full rank, with $d^2$ nonzero singular values.
-Indexing these singular values $M^V_k$ and the corresponding (normalized) right singular vectors $\v x^V_k\equiv \sum_j x^V_{kj} \ket{j}$ by an integer $k\in\set{1,2,\cdots,d^2}$, we can construct the orthonormal qudit operators
-\begin{align}
-  Q^V_k \equiv \sum_j q^V_{kj} T_j,
+Indexing these singular values $M^V_k$ and the corresponding (normalized) left singular vectors $\v x^V_k\equiv \sum_j x^V_{kj} \ket{j}$ by an integer $k\in\set{1,2,\cdots,d^2}$, we can construct the orthonormal qudit operators
+\begin{align}
+  Q^V_k \equiv \sum_j \p{q^V_{kj}}^* T_j,
   &&
@@
   &&
-  \rho_k^V \equiv \bk{Q^V_k}_\rho.
+  \rho_k^V \equiv \bk{{Q^V_k}^\dag}_\rho.
 \end{align}
@@
 \end{align}
-where
+where, using the fact that $T_j=T_j^\dag$,
 \begin{align}
@@
   \approx \sum_j q^V_{kj} \bk{T_j}_\rho
-  = \bk{Q^V_k}_\rho
+  = \bk{{Q^V_k}^\dag}_\rho
   = \rho^V_k.
@@
   &= \Bbk{\p{\tilde\rho^V_k-\rho^V_k}^*\p{\tilde\rho^V_k-\rho^V_k}} \\
-  &= \sum_{j,j'} \p{q^V_{kj}}^* q^V_{kj'}\,
+  &= \sum_{j,j'} \p{q^V_{kj}}^* q^V_{kj'} \,
   \bbk{\epsilon_j \epsilon_{j'}} \\
@@
 \end{align}
-where $M_V^{-1}$ is the right inverse of $M_V$, satisfying $M_V M_V^{-1}=1$.
+where $M_V^{-1}$ is the left inverse of $M_V$, satisfying $M_V^{-1} M_V = 1$.
 We refer to the error scale $\S_V$ as ``classical'' because the bound in Eq.~\eqref{eq:bound_eps} applies in the presence of classical sources of measurement error.
@@
 \begin{align}
-  U^\dag M_V = \sum_{\ell=0}^{d-1} \op{\ell} \otimes M_{V\ell},
-\end{align}
-where the $\p{2\ell+1}\times\abs{V}$-sized blocks are
-\begin{align}
-  M_{V\ell} \equiv \sum_{m,\v v} \ket{m} \obk{T_{\ell m}|T_{\v v\ell,0}} \bra{\v v}
-  = \sum_{m,\v v} D^\ell_{m,0}\p{\v v} \op{m}{\v v}.
+  M_V U = \sum_{\ell=0}^{d-1} \op{\ell} \otimes M_{V\ell},
+\end{align}
+where the $\abs{V}\times\p{2\ell+1}$-sized blocks are
+\begin{align}
+  M_{V\ell} \equiv \sum_{\v v,m} \ket{\v v}
+  \obk{T_{\v v\ell,0}|T_{\ell m}} \bra{m}
+  = \sum_{\v v,m} D^\ell_{0,m}\p{\v v} \op{\v v}{m}.
   \label{eq:block}
@@
 \end{align}
-where $M_{V\ell}^{-1}$ is the right inverse of $M_{V\ell}$.
+where $M_{V\ell}^{-1}$ is the left inverse of $M_{V\ell}$.
 Constructing the block $M_{V\ell}$ and computing its singular value decomposition takes at most $O(\abs{V}d^2)$ time.
@@
 We now define $\N_{V\ell}$ and $\ket{\chi_{V\ell}}$ for the sake of completion, but note that these definitions can be skipped without consequence for the remaining discussions in this paper.
-The matrix $\N_{V\ell}$ is defined in terms of the measurement matrix $M_{V\ell}$ and its right inverse $M_{V\ell}^{-1}$ by
-\begin{align}
-  \N_{V\ell} \equiv M_{V\ell} \diag\sp{M_{V\ell}^{-1}
-    \p{M_{V\ell}^{-1}}^\dag} M_{V\ell}^\dag,
+The matrix $\N_{V\ell}$ is defined in terms of the measurement matrix $M_{V\ell}$ and its left inverse $M_{V\ell}^{-1}$ by
+\begin{align}
+  \N_{V\ell} \equiv M_{V\ell}^\dag
+  \diag\sp{\p{M_{V\ell}^{-1}}^\dag M_{V\ell}^{-1}} M_{V\ell},
 \end{align}
@@
 of $\bk{T_{\v v\ell,0}}_\rho$, where the matrix elements of $T_{\ell,0}$ are provided in Eq.~\eqref{eq:trans_op}.
-\item Denoting the nonzero singular values of $M_{V\ell}$ by $M^V_{\ell k}$ and the corresponding right singular vectors by $\v x^V_{\ell k} = \sum_{\v v} x^V_{\ell k \v v} \ket{\v v}$, compute the coefficients and operators
-  \begin{align}
+\item Denoting the nonzero singular values of $M_{V\ell}$ by $M^V_{\ell k}$ and the corresponding left singular vectors by $\v x^V_{\ell k} = \sum_{\v v} x^V_{\ell k \v v} \ket{\v v}$, compute the  operators and coefficients
+  \begin{align}
+    Q_{\ell k} &\equiv \f1{M^V_{\ell k}} \sum_{\v v}
+    \p{x^V_{\ell k\v v}}^* T_{\v v\ell,0}, \\
     \tilde\rho^V_{\ell k} &\equiv \f1{M^V_{\ell k}} \sum_{\v v}
-    x^V_{\ell k\v v} \, \tilde T_{\v v\ell,0}, \\
-    Q_{\ell k} &\equiv \f1{M^V_{\ell k}} \sum_{\v v}
-    x^V_{\ell k\v v} T_{\v v\ell,0},
+    x^V_{\ell k\v v} \, \tilde T_{\v v\ell,0},
   \end{align}
@@
 Moreover, by comparing Eqs.~\eqref{eq:spin_ops} and \eqref{eq:spin_trans} we see that the polarization operators $T_{\ell m}$ transform identically to spin-$\ell$ particles under the adjoint action of the spin operators $S_\z$ and $S_\pm$.
-For any triplet of Euler angles $\v\omega=\p{\alpha,\beta,\gamma}$, we can therefore define the rotation operator
+For any triplet of angles $\v\omega=\p{\alpha,\beta,\gamma}$, we can therefore define the rotation operator
 \begin{align}
@@
 \end{align}
-and expand polarization operators in a rotated frame as
-\begin{align}
-  T_{\v\omega\ell m}
-  \equiv R\p{\v\omega}^\dag T_{\ell m} R\p{\v\omega}
-  = \sum_{n=-\ell}^\ell D_{nm}^\ell\p{\v\omega} T_{\ell n},
+and expand rotated polarization operators as
+\begin{align}
+  T_{\v\omega\ell m} \equiv
+  R\p{\v\omega} T_{\ell m} R\p{\v\omega}^\dag
+  = \sum_{n=-\ell}^\ell D_{mn}^\ell\p{\bar{\v\omega}}^* T_{\ell n},
   \label{eq:trans_rot_apndx}
 \end{align}
-where
-\begin{align}
-  D_{mn}^\ell\p{\v\omega} \equiv \bk{\ell m|R\p{\v\omega}|\ell n}
-  = \obk{T_{\ell m} | R\p{\v\omega}^\dag
-    \otimes R\p{\v\omega}^{\t{T}} |T_{\ell n}}
-\end{align}
-are matrix elements of the rotation operator $R\p{\v\omega}$ for spin-$\ell$ particles, and $R\p{\v\omega}^{\t{T}}$ is the transpose of $R\p{\v\omega}$.
-
-Implicitly associating the Euler angle doublet $\v v = \p{\alpha,\beta}$ with the triplet $\p{\alpha,\beta,0}$, the transformation rules in Eq.~\eqref{eq:trans_rot_apndx} imply that we can expand the phase-space representation of $T_{\ell m}$ as
+where $\bar{\v\omega}=\p{\gamma,\beta,\alpha}$ is the reversal of $\v\omega$, and
+\begin{align}
+  D_{mn}^\ell\p{\bar{\v\omega}}
+  \equiv \bk{\ell m|R\p{\bar{\v\omega}}|\ell n}
+  = \obk{T_{\ell n} | R\p{\v\omega}
+    \otimes R\p{\v\omega}^* |T_{\ell m}}^*
+  = \obk{T_{\ell m} | R\p{-\bar{\v\omega}}
+    \otimes R\p{-\bar{\v\omega}}^* |T_{\ell n}}
+\end{align}
+are matrix elements of the rotation operator $R\p{\v\omega}$ for spin-$\ell$ particles.
+
+For any angle doublet $\v v=\p{\alpha,\beta}$, we define $R\p{\v v}\equiv R\p{\alpha,\beta,0}$ and $D^\ell_{mn}\p{\v v}=D^\ell_{mn}\p{0,\beta,\alpha}$ for shorthand.
+The transformation rules in Eq.~\eqref{eq:trans_rot_apndx} imply that we can expand the phase-space representation of $T_{\ell m}$ as
 \begin{align}
@@
 \begin{align}
-  M_V \equiv \sum_{\v v,\ell} \oket{T_{\v v\ell,0}} \bra{\v v\ell},
+  M_V \equiv \sum_{\v v,\ell} \ket{\v v\ell} \obra{T_{\v v\ell,0}},
 \end{align}
@@
 \begin{align}
-  U^\dag M_V = \sum_\ell \op{\ell} \otimes M_{V\ell},
+  M_V U = \sum_\ell \op{\ell} \otimes M_{V\ell},
   &&
@@
   &&
-  M_{V\ell} = \sum_{m,\v v} D^\ell_{m,0}\p{\v v} \op{m}{\v v},
+  M_{V\ell} = \sum_{m,\v v} D^\ell_{m,0}\p{\v v} \op{\v v}{m},
 \end{align}
 where $D^\ell_{mn}\p{\v v}\equiv\bk{\ell m|R\p{\v v}|\ell n}$ is a (Wigner) rotation matrix element for a spin-$\ell$ particle.
-The block-diagonal structure of $M_V$ allows us to index its singular values $M^V_{\ell m}$ and corresponding (normalized) right singular vectors $\v x^V_{\ell m} = \sum_i x^V_{\ell m i} \ket{\v v_i}$ by the indices $\p{\ell,m}$, where the integer $\abs{m}\le\ell$.
+The block-diagonal structure of $M_V$ allows us to index its singular values $M^V_{\ell m}$ and corresponding (normalized) left singular vectors $\v x^V_{\ell m} = \sum_i x^V_{\ell m i} \ket{\v v_i}$ by the indices $\p{\ell,m}$, where the integer $\abs{m}\le\ell$.
 These singular vectors and values define the orthonormal operators
 \begin{align}
-  Q^V_{\ell m} \equiv \sum_i q^V_{\ell mi} T_{i\ell},
+  Q^V_{\ell m} \equiv \sum_i \p{q^V_{\ell mi}}^* T_{i\ell},
   &&
@@
 \begin{align}
-  \rho = \sum_{\ell,m} \bk{Q^V_{\ell m}}_\rho Q^V_{\ell m},
+  \rho = \sum_{\ell,m} \bk{{Q^V_{\ell m}}^\dag}_\rho Q^V_{\ell m},
 \end{align}
@@
 \end{align}
+Here $M_{V\ell}^{-1}$ is the left inverse of $M_{V\ell}$.
 The fact that $\bbk{\epsilon_{i\ell}^2}<1/2n$ also implies that
@@
   = \sum_{\ell,m,i,i'} \p{q^V_{\ell mi}}^* q^V_{\ell mi'} \bbk{\epsilon_{i\ell}\epsilon_{i'\ell}}
-  = \f1n \sum_{\ell,m,i} \abs{\tilde{\v q}_{\ell i}}^2 \cov_\rho\p{T_{i\ell},T_{i\ell}},
+  = \f1n \sum_{\ell,i} \abs{\tilde{\v q}_{\ell i}}^2 \cov_\rho\p{T_{i\ell},T_{i\ell}},
   &&
@@
 \end{align}
-where $\tilde{\v q}_{\ell i} = \sum_m q_{\ell mi} \ket{m}$, and we used the fact that $\bbk{\epsilon_{i\ell}\epsilon_{i'\ell}} = \delta_{ii'}\times\cov_\rho\p{T_{i\ell},T_{i\ell}}/n$.
-Identifying the singular value decomposition $M_{V\ell} = U_{V\ell} \Sigma_{V\ell} W_{V\ell}^\dag$, we then we observe that vector $\tilde{\v q}_{\ell i}$ is essentially the $i$-th row of $\Sigma_{V\ell}^{-1} W_{V\ell}^\dag$, which is to say that $\tilde{\v q}_{\ell i} = \Sigma_{V\ell}^{-1} W_{V\ell}^\dag \ket{\v v_i}$.
-We can therefore simplify
+where $\tilde{\v q}_{\ell i} = \sum_m \p{q_{\ell mi}}^* \ket{m}$, and we used the fact that $\bbk{\epsilon_{i\ell}\epsilon_{i'\ell}} = \delta_{ii'}\times\cov_\rho\p{T_{i\ell},T_{i\ell}}/n$.
+Identifying the singular value decomposition $M_{V\ell} = U_{V\ell} \Sigma_{V\ell} W_{V\ell}^\dag$, we then we observe that $\tilde{\v q}_{\ell i} = \Sigma_{V\ell}^{-1} U_{V\ell}^\dag \ket{\v v_i}$, which allows us to simplify
 \begin{align}
   \abs{\tilde{\v q}_{\ell i}}^2
-  = \bk{\v v_i | W_{V\ell} \Sigma_{V\ell}^{-2} W_{V\ell}^\dag | \v v_i}
-  = \bk{\v v_i| M_{V\ell}^{-1} \p{M_{V\ell}^{-1}}^\dag | \v v_i}.
+  = \bk{\v v_i | U_{V\ell} \Sigma_{V\ell}^{-2} U_{V\ell}^\dag | \v v_i}
+  = \bk{\v v_i| \p{M_{V\ell}^{-1}}^\dag M_{V\ell}^{-1} | \v v_i}.
 \end{align}
@@
   = \cov_\rho\p{T_{i\ell}^\dag,T_{i\ell}}
-  = \sum_{m,m'} D^\ell_{m,0}\p{\v v_i}^* D^\ell_{m',0}\p{\v v_i}
+  = \sum_{m,m'} D^\ell_{0,m}\p{\v v_i} D^\ell_{0,m'}\p{\v v_i}^*
   \cov_\rho\p{T_{\ell m}^\dag, T_{\ell m'}},
@@
   \E_V\p{\rho}^2
-  = \f1n \sum_{i,\ell,m,m'}
-  D^\ell_{m',0}\p{\v v_i} \abs{\tilde{\v q}_{\ell i}}^2
-  D^\ell_{m,0}\p{\v v_i}^*
+  = \f1n \sum_{\ell,i,m,m'}
+  D^\ell_{0,m'}\p{\v v_i}^* \abs{\tilde{\v q}_{\ell i}}^2
+  D^\ell_{0,m}\p{\v v_i}
   \cov_\rho\p{T_{\ell m}^\dag, T_{\ell m'}}.
@@
 \begin{align}
-  \N_{V\ell} \equiv M_{V\ell} \diag\sp{M_{V\ell}^{-1}
-    \p{M_{V\ell}^{-1}}^\dag} M_{V\ell}^\dag,
+  \N_{V\ell} \equiv M_{V\ell}^\dag \diag\sp{\p{M_{V\ell}^{-1}}^\dag
+    M_{V\ell}^{-1}} M_{V\ell},
 \end{align}

### 2020-12-04 sun_phases/scripts/qudit_tomo_methods.py

@@ -6,6 +6,8 @@ import numpy as np
 
 import wigner.lib, py3nj
 
+_update = True
+
 ##########################################################################################
 # general simulation methods
 
@@ -55,20 +57,19 @@ def compute_batch(function, args):
     return { arg : value for arg, value in zip(args_list, values) }
 
 ##########################################################################################
-# methods to compute a "measurement matrix" of all D^L_{m,0}(v) with a fixed degree L
-# where: D^L_{mn}(v) = <Lm |R(v)|Ln> is a (Wigner) rotation matrix element
-#        v = (alpha,beta) is a point on the sphere at azimuth/polar angles (alpha,beta)
-#        R(v) is an Euler-angle rotation operator: exp(-i alpha S_z) * exp(-i beta S_y)
-#        S_z and S_y are respectively spin-z and spin_y operators
+# methods to compute a "measurement matrix" of all D^L_{0,m}(v) with a fixed degree L
+# where: v = (alpha,beta) is a point on the sphere at azimuth/polar angles (alpha,beta)
+#        D^L_{mn}(v) = <Lm| e^{-i beta S_y} * e^{-i alpha S_z} |Ln>
+#        S_y and S_z are respectively spin-y and spin-z operators
 #        |Lm> is a state of a spin-L particle with spin projection m onto the z axis
 #
-# the basic idea behind our methods is to decompose D^L_{m,0}(alpha,beta) into:
-#     e^{-i alpha S_z} * exp(+i pi/2 S_y) * e^{-i beta S_z} * exp(-i pi/2 S_y) |L,0>
+# the basic idea behind our methods is to decompose D^L_{0,m}(v) into:
+#     <L,0| exp(+i pi/2 S_y) * e^{-i beta S_z} * exp(-i pi/2 S_y) * e^{-i alpha S_z}
 
 # methods to pre-compute spin values and pi/2 rotation matrices exp(-i pi/2 S_y)
 def _spin_vals(LL):
     return np.arange(-LL, LL+1)
-def _rot_z_to_x(LL, update = True):
+def _rot_z_to_x(LL, update = _update):
     if update:
         print("rot:", LL)
         sys.stdout.flush()
@@ -83,21 +84,20 @@ def get_meas_methods(max_dim):
[…]

### 2020-12-10 sun_phases/spin_qudit_tomo.tex
added exact 'classical' reconstruction error, updated some todos, added citation for pending codes

@@
   We then provide algorithms with $O\p{rd^3}$ serial runtime, parallelizable down to $O\p{rd^2}$, for (i) computing a priori upper bounds on the expected error with which spin projection measurements along $r$ given axes can reconstruct an unknown qudit state, and (ii) estimating a posteriori the statistical error in a reconstructed state.
-  Our algorithms motivate a simple randomized tomography protocol, for which we numerically find that using more measurement axes can yield substantial benefits that plateau after $r\approx3d$.
+  Our algorithms motivate a simple randomized tomography protocol, for which we find that using more measurement axes can yield substantial benefits that plateau after $r\approx3d$.
 \end{abstract}
@@
 These techniques are typically developed in a general, information-theoretic setting, and make minimal assumptions about the physical medium of a quantum state.
-As a consequence, even well-established techniques can nonetheless be ill-suited for physical platforms with unique or limited capabilities.
-
-Due to advancements in experimental capabilities to address nuclear spin states (hyperfine levels) in ultracold atomic systems \cite{daley2011quantum, mischuck2012control, smith2013quantum, cazalilla2014ultracold}, as well as developments in the control of ultracold molecular systems \cite{bohn2017cold, zeppenfeld2012sisyphus, kozyryev2017sisyphus, puri2017synthesis, wu2017cryofuge, marco2019degenerate, liu2019molecular, anderegg2019optical, chou2020frequencycomb, lin2020quantum}, a particular setting of growing interest is the spin qudit, or a multilevel quantum angular momentum degree of freedom.
+As a consequence, even well-established techniques can be ill-suited for physical platforms with unique or limited capabilities.
+
+% todo: review / revisit tomography examples above
+
+Due to advancements in experimental capabilities to address nuclear spin states (i.e.~hyperfine levels) in ultracold atomic systems \cite{daley2011quantum, mischuck2012control, smith2013quantum, cazalilla2014ultracold}, as well as developments in the control of ultracold molecular systems \cite{bohn2017cold, zeppenfeld2012sisyphus, kozyryev2017sisyphus, puri2017synthesis, wu2017cryofuge, marco2019degenerate, liu2019molecular, anderegg2019optical, chou2020frequencycomb, lin2020quantum}, a particular setting of growing interest is the spin qudit, or a multilevel quantum angular momentum degree of freedom.
 Spin qudits can provide advantages over their qubit counterparts for quantum sensing \cite{hemmer2018squeezing, evrard2019enhanced}, enable quantum simulations of SU($d$) magnetism \cite{cazalilla2014ultracold, banerjee2013atomic, zhang2014spectroscopic, scazza2014observation, goban2018emergence}, and offer unique capabilities for quantum error correction \cite{albert2020robust, gross2020encoding}.
 In all cases, quantum state tomography is necessary to take full advantage of a spin qudit.
+
+% todo: explain the role of tomography in sensing and simulation?
 
@@
 In particular, this mapping allows us to bound the minimum number of measurement axes necessary to perform full tomography on a $d$-level spin qudit, $r_d^{\t{min}}$, by $2d-1\le r_d^{\t{min}}\le d^2$.
-In Section \ref{sec:error} we provide (a priori) upper bounds and (a posteriori) estimates of the statistical error in a qudit state reconstructed from measurements of spin projection along a given set of $r$ measurement axes.
+In Section \ref{sec:error} we provide a priori upper bounds and a posteriori estimates of the statistical error in a qudit state reconstructed from measurements of spin projection along a given set of $r$ measurement axes.
 The capability to determine upper bounds on reconstruction error a priori motivates a simple randomized tomography protocol that we outline in Section \ref{sec:protocol}, and for which we numerically find that using more measurement axes yields substantial benefits that plateau after $r\approx3d$.
-
-% todo: comment on code availability
+To facilitate the use of our protocols, we make all of our codes publicly available at Ref.~\cite{codes}.
+
+% todo: add figure for spin measurements?
 
@@
 are (Wigner) rotation matrix elements.
-Throughout this work, we will primarily consider rotations of the sphere that take the north pole to a point $\v v=\p{\alpha,\beta}$ at azimuthal angle $\alpha$ and polar angle $\beta$.
+For reasons that will become clear shortly, throughout this work we will primarily consider rotations of the sphere that take the north pole to a point $\v v=\p{\alpha,\beta}$ at azimuthal angle $\alpha$ and polar angle $\beta$.
 For ease of notation, we therefore define $R\p{\v v} \equiv R\p{\alpha,\beta,0}$, $T_{\v v\ell m} \equiv T_{\p{\alpha,\beta,0},\ell m}$, and $D^\ell_{mn}\p{\v v} \equiv D^\ell_{mn}\p{0,\beta,\alpha}$.
@@
 
-% todo: add figure for signal recovery problem?
+% todo: add figure for signal recovery problem
 
@@
 
-The quantum error scale $\epsilon_V$ provides a pessimistic upper bound on statistical error, which can be calculated without prior knowledge of the true qudit state $\rho$.
-The actual shot-noise-limited error in the reconstruction $\tilde\rho_V$ of a particular state $\rho$ may be considerably smaller, and depends on $\rho$ itself.
-The exact mean squared distance between $\tilde\rho_V$ and $\rho$, derived in Appendix \ref{sec:exact}, can be written in the form
-\begin{align}
-  \E_V\p{\rho}^2
-  = \f1n \sum_\ell \sp{\bk{\chi_{V\ell}|\rho_\ell}
+The error scales $\S_V$ and $\epsilon_V$ provide pessimistic upper bounds on statistical error, which can be calculated without prior knowledge of the true qudit state $\rho$.
+The actual error in the reconstruction $\tilde\rho_V$ of a particular state $\rho$ may be considerably smaller, and may depend on $\rho$ itself.
+Written out in full, the mean squared distance between $\tilde\rho_V$ and $\rho$ is (see Appendix \ref{sec:bound})
+\begin{align}
+  \E_V\p{\rho}^2 = \sum_{\v v,\v w,\ell}
+  \bk{\v v|\p{M_{V\ell}^{-1}}^\dag M_{V\ell}^{-1}|\v w}
+  \bbk{\epsilon_{\v v\ell}\epsilon_{\v w\ell}},
+  \label{eq:error_cls}
+\end{align}
+where $\bbk{\epsilon_{\v v\ell}\epsilon_{\v w\ell}}$ is the covariance between errors $\epsilon_{\v v\ell}$ in the empirical estimates $\tilde T_{\v v\ell,0}$ of $\bk{T_{\v v\ell,0}}_\rho$.
+These covariances are generally determined by the sources of measurement error in any given experiment, but will typically satisfy $\bbk{\epsilon_{\v v\ell}\epsilon_{\v w\ell}} = \delta_{\v v\v w}\bbk{\epsilon_{\v v\ell}^2}$.
+If measurement error is limited by shot noise, then (see Appendix \ref{sec:exact})
+\begin{align}
+  \E_V\p{\rho}^2 \stackrel{\t{SNL}}{=} \f1n \sum_\ell
+  \sp{\bk{\chi_{V\ell}|\rho_\ell}
     - \bk{\rho_\ell|\N_{V\ell}|\rho_\ell}},
-  \label{eq:error}
-\end{align}
-where $\ket{\rho_\ell} \equiv \sum_m \rho_{\ell m}\ket{m}$ is a vector of the degree-$\ell$ polarization operator components $\rho_{\ell m}$ of $\rho$, defined in Eq.~\eqref{eq:trans_state}, and the matrix $\N_{V\ell}$ and vector $\ket{\chi_{V\ell}}$ are defined below.
-While the true error in $\tilde\rho_V$ cannot be known exactly without knowing $\rho$, this error can be estimated a posteriori by $\E_V\p{\tilde\rho_V} \approx \E_V\p{\rho}$.
-After constructing an estimate $\tilde\rho_V$ of $\rho$, the complexity of computing $\E_V\p{\tilde\rho_V}$ is the same as that of $\S_V$ and $\epsilon_V$ (see Figure \ref{fig:times}).
+  \label{eq:error_qnt}
+\end{align}
+where $\stackrel{\t{SNL}}{=}$ indicates equality in the ``shot-noise-limited'' regime; $\ket{\rho_\ell} \equiv \sum_m \rho_{\ell m}\ket{m}$ is a vector of the scalar polarization operator components $\rho_{\ell m}$ of $\rho$, defined in Eq.~\eqref{eq:trans_state}; and the matrix $\N_{V\ell}$ and vector $\ket{\chi_{V\ell}}$ are defined below.
+While the true shot-noise-limited error in $\tilde\rho_V$ cannot be known exactly without knowing $\rho$, this error can be estimated a posteriori by $\E_V\p{\rho}\approx\E_V\p{\tilde\rho_V}$.
+After constructing an estimate $\tilde\rho_V$ of $\rho$, the complexity of computing the error $\E_V\p{\tilde\rho_V}$ from Eq.~\eqref{eq:error_qnt} is the same as that of computing $\S_V$ or $\epsilon_V$ (see Figure \ref{fig:times}).
 
 We now define $\N_{V\ell}$ and $\ket{\chi_{V\ell}}$ for the sake of completion, but note that these definitions can be skipped without consequence for the remaining discussions in this paper.
-The matrix $\N_{V\ell}$ is defined in terms of the measurement matrix $M_{V\ell}$ and its left inverse $M_{V\ell}^{-1}$ by
+The matrix $\N_{V\ell}$ is
 \begin{align}
@@
 \end{align}
-where $\diag\sp{X}$ sets all off-diagonal parts of $X$ to zero.
+where $\diag\sp{X}$ sets the off-diagonal parts of $X$ to zero.
 The vector $\ket{\chi_{V\ell}}\equiv\sum_m \chi^V_{\ell m} \ket{m}$ is defined by
@@
 
+% todo: look up and cite randomized protocols
+% see JQI / Jake Taylor / Mohammad Hafezi
+
 The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following protocol for spin qudit tomography:
@@
   Note that computing the error scale $\epsilon_V$ requires, for each $\ell\in\set{0,1,\cdots,d-1}$, constructing the measurement matrix $M_{V\ell}$ in Eq.~\eqref{eq:block} and computing its singular value decomposition.
-  Save all measurement matrix data associated with the final measurement axes $V$ for later use. % todo: cite codes
+  Save all measurement matrix data associated with the final measurement axes $V$ for later use.
 \item For each axis $\v v\in V$, make $n$ measurements of spin projection, and set $\tilde\Pi_{\v v\mu}\approx\bk{\Pi_{\v v\mu}}_\rho$ to the fraction of times in which the measurement outcome was $\mu$.
@@
 \end{enumerate}
-If reconstruction error is dominated by shot noise, then the total error in $\tilde\rho_V$, or its root-mean-square distance from $\rho$, is approximately $\E_V\p{\tilde\rho_V}\approx\E_V\p{\rho}$, which can be computed from Eq.~\eqref{eq:error}.
-% todo: cite codes
+Th expected reconstruction error in $\tilde\rho_V$, or its root-mean-square distance from $\rho$, is provided by Eq.~\eqref{eq:error_cls}.
+If measurement error is shot-noise-limited, then the error in $\tilde\rho_V$ is approximately $\E_V\p{\rho}\approx\E_V\p{\tilde\rho_V}$ and can be computed from Eq.~\eqref{eq:error_qnt}.
 If $\tilde\rho_V$ has negative eigenvalues, its distance from $\rho$ can be reduced with maximum-likelihood corrections \cite{smolin2012efficient}.
@@
 % todo: include table of good measurement axes for specific values of d
-% Yb (5/2), Cr (3), Sr (9/2), Dy (15/2?), Er (19/2)
+% Yb-173 (6), Cr (7), Sr-87 (10), Er-168 (13), Dy (17), Er-167 (20)
 

### 2020-12-11 sun_phases/spin_qudit_tomo.tex
added some missing references, changed documentclass

@@
-\documentclass[notitlepage,twocolumn]{revtex4-2}
+\documentclass[a4paper,twocolumn,unpublished]{quantumarticle}
+\pdfoutput=1
+
+%%% properly deal with different font encodings
+\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
+
+%%% bibliography
+\usepackage[backend=bibtex,style=phys,doi=false]{biblatex}
+\addbibresource{\jobname.bib}
 
@@
 
-%%% properly deal with different font encodings
-\usepackage[utf8]{inputenc}
-\usepackage[T1]{fontenc}
+% use alphanumeric tags for footnotes
+\renewcommand*{\thefootnote}{\alph{footnote}}
 
@@
 \newcommand{\bk}{\Braket} % shorthand for braket notation
-\renewcommand{\v}{\bm} % bold vectors
 \newcommand{\uv}[1]{\bm{\hat{#1}}} % unit vectors
@@
 
-\newcommand{\JILA}{JILA, National Institute of Standards and Technology and
-  University of Colorado, 440 UCB, Boulder, Colorado 80309, USA}
-\newcommand{\CTQM}{Center for Theory of Quantum Matter, University of Colorado, Boulder, CO, 80309, USA}
-
-\newcommand{\thetitle}{Spin qudit tomography}
-
-\title{\thetitle}
+\title{Spin qudit tomography}
 \author{Michael A.~Perlin}
 \email{mika.perlin@gmail.com}
+\orcid{https://orcid.org/0000-0002-9316-1596}
 \author{Diego Barberena}
 \author{Ana Maria Rey}
-\affiliation{\JILA}
-\affiliation{\CTQM}
-\date{\today}
+\orcid{http://orcid.org/0000-0001-7176-9413}
+\affiliation{JILA, National Institute of Standards and Technology and University of Colorado, 440 UCB, Boulder, Colorado 80309, USA}
+\affiliation{Center for Theory of Quantum Matter, University of Colorado, Boulder, CO, 80309, USA}
+\date{10 December 2020}
 
@@
 Quantum state tomography, the task of reconstructing a quantum state by collecting and processing measurement data, is an essential primitive for quantum sensing, quantum simulation, and quantum information processing.
-The central importance of quantum state tomography has led to the development of a variety of techniques based on compressed sensing \cite{gross2010quantum}, matrix product states \cite{cramer2010efficient}, maximum-likelihood estimation \cite{smolin2012efficient}, Bayesian inference \cite{huszar2012adaptive}, and neural networks \cite{torlai2018neuralnetwork}, among many others \cite{christandl2012reliable, qi2013quantum, opatrny1997leastsquares}.
-These techniques are typically developed in a general, information-theoretic setting, and make minimal assumptions about the physical medium of a quantum state.
+The central importance of quantum state tomography has led to the development of techniques based on least-squares inversion \cite{opatrny1997leastsquares}, linear regression \cite{qi2013quantum}, maximum likelihood estimation \cite{teo2011quantumstate, smolin2012efficient}, Bayesian inference \cite{huszar2012adaptive, ferrie2014quantum, granade2016practical}, compressed sensing \cite{gross2010quantum, kalev2015quantum}, neural networks \cite{torlai2018neuralnetwork}, among others too numerous to exhaustively list here.
+% todo: fix citation for teo2011
+These techniques are typically developed in a general information-theoretic setting, and make minimal assumptions about the physical medium of a quantum state.
 As a consequence, even well-established techniques can be ill-suited for physical platforms with unique or limited capabilities.
-
-% todo: review / revisit tomography examples above
 
@@
 The defining property of a spin qudit, distinguishing it from other qudits, is that it describes an angular momentum degree of freedom, which has specific implications for how a spin qudit should transform under the group SO(3) of rotations in 3D space.
-Due to the central importance of these transformation rules for a spin qudit, we seek a basis of operators that transform nicely under 3D rotations \footnote{Technically speaking, we seek a basis of operators that transform as an irreducible representation of SO(3).}.
+Due to the central importance of these transformation rules for a spin qudit, we seek a basis of operators that transform nicely under 3D rotations\footnote{Technically speaking, we seek a basis of operators that transform as an irreducible representation of SO(3).}.
 One such basis is that of the {\it polarization operators} \cite{kryszewski2006positivity, bertlmann2008bloch}, defined by
@@
 Moreover, the degree-$\ell$ polarization operators $T_{\ell m}$ transform similarly to spin-$\ell$ particles and spherical harmonics $Y_{\ell m}$ under 3D rotations (see Appendix \ref{sec:rotations}).
-Specifically, for any triplet of angles $\v\omega=\p{\alpha,\beta,\gamma}$, we can therefore define the rotation operator
-\begin{align}
-  R\p{\v\omega} \equiv e^{-\i\alpha S_\z} e^{-\i\beta S_\y} e^{-\i\gamma S_\z},
+Specifically, for any triplet of angles $\bm\omega=\p{\alpha,\beta,\gamma}$, we can therefore define the rotation operator
+\begin{align}
+  R\p{\bm\omega} \equiv e^{-\i\alpha S_\z} e^{-\i\beta S_\y} e^{-\i\gamma S_\z},
 \end{align}
@@
 \begin{align}
-  T_{\v\omega\ell m} \equiv
-  R\p{\v\omega} T_{\ell m} R\p{\v\omega}^\dag
-  = \sum_{n=-\ell}^\ell D_{mn}^\ell\p{\bar{\v\omega}}^* T_{\ell n},
+  T_{\bm\omega\ell m} \equiv
+  R\p{\bm\omega} T_{\ell m} R\p{\bm\omega}^\dag
+  = \sum_{n=-\ell}^\ell D_{mn}^\ell\p{\bar{\bm\omega}}^* T_{\ell n},
   \label{eq:trans_rot}
 \end{align}
-where $\bar{\v\omega}=\p{\gamma,\beta,\alpha}$ is the reversal of $\v\omega$, and
-\begin{align}
-  D_{mn}^\ell\p{\bar{\v\omega}}
-  \equiv \bk{\ell m|R\p{\bar{\v\omega}}|\ell n}
+where $\bar{\bm\omega}=\p{\gamma,\beta,\alpha}$ is the reversal of $\bm\omega$, and
+\begin{align}
+  D_{mn}^\ell\p{\bar{\bm\omega}}
+  \equiv \bk{\ell m|R\p{\bar{\bm\omega}}|\ell n}
 \end{align}
 are (Wigner) rotation matrix elements.
-For reasons that will become clear shortly, throughout this work we will primarily consider rotations of the sphere that take the north pole to a point $\v v=\p{\alpha,\beta}$ at azimuthal angle $\alpha$ and polar angle $\beta$.
-For ease of notation, we therefore define $R\p{\v v} \equiv R\p{\alpha,\beta,0}$, $T_{\v v\ell m} \equiv T_{\p{\alpha,\beta,0},\ell m}$, and $D^\ell_{mn}\p{\v v} \equiv D^\ell_{mn}\p{0,\beta,\alpha}$.
+For reasons that will become clear shortly, throughout this work we will primarily consider rotations of the sphere that take the north pole to a point $\bm v=\p{\alpha,\beta}$ at azimuthal angle $\alpha$ and polar angle $\beta$.
+For ease of notation, we therefore define $R\p{\bm v} \equiv R\p{\alpha,\beta,0}$, $T_{\bm v\ell m} \equiv T_{\p{\alpha,\beta,0},\ell m}$, and $D^\ell_{mn}\p{\bm v} \equiv D^\ell_{mn}\p{0,\beta,\alpha}$.
 
@@
 In fact, the phase-space representation of $T_{\ell m}$ is proportional to $Y_{\ell m}$.
-The phase-space representation of a spin qudit operator $X$ assigns, to each point $\v v$ on the sphere, the complex number
-\begin{align}
-  X^{\t{PS}}\p{\v v} \equiv \bk{s_{\v v}|X|s_{\v v}},
-\end{align}
-where $\ket{s_{\v v}}\equiv R\p{\v v}\ket{s}$ is the state of a spin qudit polarized along $\v v$.
-This representation is faithful in the sense that $X$ is uniquely determined by the phase-space values $X^{\t{PS}}\p{\v v}$ at all points $\v v$ on the sphere.
+The phase-space representation of a spin qudit operator $X$ assigns, to each point $\bm v$ on the sphere, the complex number
+\begin{align}
+  X^{\t{PS}}\p{\bm v} \equiv \bk{s_{\bm v}|X|s_{\bm v}},
+\end{align}
+where $\ket{s_{\bm v}}\equiv R\p{\bm v}\ket{s}$ is the state of a spin qudit polarized along $\bm v$.
+This representation is faithful in the sense that $X$ is uniquely determined by the phase-space values $X^{\t{PS}}\p{\bm v}$ at all points $\bm v$ on the sphere.
 The transformation rules for polarization operators in Eq.~\eqref{eq:trans_rot}, together with the fact that $\bk{s|T_{\ell m}|s}=0$ unless $m=0$, suffice to show that (see Appendix \ref{sec:rotations})
 \begin{align}
-  T_{\ell m}^{\t{PS}}\p{\v v} = c_\ell Y_{\ell m}\p{\v v},
+  T_{\ell m}^{\t{PS}}\p{\bm v} = c_\ell Y_{\ell m}\p{\bm v},
 \end{align}
@@
 Performing tomography on an unknown qudit state $\rho$ is therefore equivalent to reconstructing the unknown distribution $\rho^{\t{PS}}$ on the sphere.
-In principle, the representation $\rho^{\t{PS}}$ of a finite-dimensional qudit state $\rho$ can be reconstructed from the values $\rho^{\t{PS}}\p{\v v}=\bk{s_{\v v}|\rho|s_{\v v}}$ at a finite number of points $\v v$.
-In practice, the value $\bk{s_{\v v}|\rho|s_{\v v}}$ is determined by measuring spin projection along $\v v$, which also provides measurement data on all spin projections $\bk{\mu_{\v v}|\rho|\mu_{\v v}}$ with $\mu\in\set{s,s-1,\cdots,-s}$ and $\ket{\mu_{\v v}}\equiv R\p{\v v}\ket{\mu}$; one would like to make use of this additional data as well.
-We clarify the connection between the quantum problem of reconstructing $\rho$ from spin projection measurements and the classical problem of reconstructing $\rho^{\t{PS}}$ from its values $\rho^{\t{PS}}\p{\v v}$ in the following section.
-In addition to providing a new perspective on the problem of spin qudit tomography, this connection will allow us to place upper bounds on the minimal number of measurement axes $\v v$ necessary for full recovery of an unknown state.
+In principle, the representation $\rho^{\t{PS}}$ of a finite-dimensional qudit state $\rho$ can be reconstructed from the values $\rho^{\t{PS}}\p{\bm v}=\bk{s_{\bm v}|\rho|s_{\bm v}}$ at a finite number of points $\bm v$.
+In practice, the value $\bk{s_{\bm v}|\rho|s_{\bm v}}$ is determined by measuring spin projection along $\bm v$, which also provides measurement data on all spin projections $\bk{\mu_{\bm v}|\rho|\mu_{\bm v}}$ with $\mu\in\set{s,s-1,\cdots,-s}$ and $\ket{\mu_{\bm v}}\equiv R\p{\bm v}\ket{\mu}$; one would like to make use of this additional data as well.
+We clarify the connection between the quantum problem of reconstructing $\rho$ from spin projection measurements and the classical problem of reconstructing $\rho^{\t{PS}}$ from its values $\rho^{\t{PS}}\p{\bm v}$ in the following section.
+In addition to providing a new perspective on the problem of spin qudit tomography, this connection will allow us to place upper bounds on the minimal number of measurement axes $\bm v$ necessary for full recovery of an unknown state.
 
@@
 Our goal is to reconstruct an arbitrary state $\rho$ of a spin qudit from measurements of spin projection onto different quantization axes.
-We are thus nominally restricted to measuring projectors $\Pi_{\v v\mu} \equiv \op{\mu_{\v v}}$, where $\ket{\mu_{\v v}}\equiv R\p{\v v}\ket{\mu}$ is a state with spin projection $\mu$ onto the measurement axis $\v v$.
-For any fixed axis $\v v$, the sets $\set{\Pi_{\v v\mu}}$ and $\set{T_{\v v\ell,0}}$ are both complete bases for the space of operators that are diagonal in the basis $\set{\ket{\mu_{\v v}}}$.
-Measuring the projectors $\set{\Pi_{\v v\mu}}$ is therefore equivalent to measuring the polarization operators $\set{T_{\v v\ell,0}}$, and provides data on the expectation values $\bk{T_{\v v\ell,0}}_\rho$.
-
-In order to reconstruct an arbitrary density operator $\rho$ from the expectation values $\bk{T_{\v v\ell,0}}_\rho$, we essentially need to to find a set of coefficients $C_{\ell mk}\p{\v v}$ that would allow us to recover any matrix element $\rho_{\ell m}$ of $\rho$ through
+We are thus nominally restricted to measuring projectors $\Pi_{\bm v\mu} \equiv \op{\mu_{\bm v}}$, where $\ket{\mu_{\bm v}}\equiv R\p{\bm v}\ket{\mu}$ is a state with spin projection $\mu$ onto the measurement axis $\bm v$.
+For any fixed axis $\bm v$, the sets $\set{\Pi_{\bm v\mu}}$ and $\set{T_{\bm v\ell,0}}$ are both complete bases for the space of operators that are diagonal in the basis $\set{\ket{\mu_{\bm v}}}$.
+Measuring the projectors $\set{\Pi_{\bm v\mu}}$ is therefore equivalent to measuring the polarization operators $\set{T_{\bm v\ell,0}}$, and provides data on the expectation values $\bk{T_{\bm v\ell,0}}_\rho$.
+
+In order to reconstruct an arbitrary density operator $\rho$ from the expectation values $\bk{T_{\bm v\ell,0}}_\rho$, we essentially need to to find a set of coefficients $C_{\ell mk}\p{\bm v}$ that would allow us to recover any matrix element $\rho_{\ell m}$ of $\rho$ through
 \begin{align}
   \rho_{\ell m}^* = \bk{T_{\ell m}}_\rho
-  = \sum_{\v v,k} C_{\ell mk}\p{\v v} \bk{T_{\v vk,0}}_\rho.
+  = \sum_{\bm v,k} C_{\ell mk}\p{\bm v} \bk{T_{\bm vk,0}}_\rho.
   \label{eq:state_recon}
 \end{align}
-Expanding the rotated polarization operators $T_{\v vk,0}$ into a sum of un-rotated polarization operators, according to Eq.~\eqref{eq:trans_rot}, we find that the recovery condition in Eq.~\eqref{eq:state_recon} is satisfied when
+Expanding the rotated polarization operators $T_{\bm vk,0}$ into a sum of un-rotated polarization operators, according to Eq.~\eqref{eq:trans_rot}, we find that the recovery condition in Eq.~\eqref{eq:state_recon} is satisfied when
 \begin{align}
   T_{\ell m}
-  = \sum_{\v v,k,n} C_{\ell mk}\p{\v v} D^k_{0,n}\p{\v v}^* T_{kn}.
-\end{align}
-Orthogonality of the polarization operators then implies the decomposition $C_{\ell mk}\p{\v v}=\delta_{\ell k}C_{\ell m}\p{\v v}$, and in turn
-\begin{align}
-  \sum_{\v v} C_{\ell m}\p{\v v} D^\ell_{0,n}\p{\v v}^*
+  = \sum_{\bm v,k,n} C_{\ell mk}\p{\bm v} D^k_{0,n}\p{\bm v}^* T_{kn}.
+\end{align}
+Orthogonality of the polarization operators then implies the decomposition $C_{\ell mk}\p{\bm v}=\delta_{\ell k}C_{\ell m}\p{\bm v}$, and in turn
+\begin{align}
+  \sum_{\bm v} C_{\ell m}\p{\bm v} D^\ell_{0,n}\p{\bm v}^*
   = \delta_{mn}
@@
 
-In fact, the problem of finding suitable axes $V$ and coefficients $C_{\ell m}\p{\v v}$ to satisfy Eq.~\eqref{eq:tomo_recovery} can be mapped onto the well-studied problem of signal recovery on the sphere \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
+In fact, the problem of finding suitable axes $V$ and coefficients $C_{\ell m}\p{\bm v}$ to satisfy Eq.~\eqref{eq:tomo_recovery} can be mapped onto the well-studied problem of signal recovery on the sphere \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
 The signal recovery problem can be stated as follows: given a square-integrable function $f$ on the sphere, with the spherical harmonic expansion
 \begin{align}
-  f\p{\v v} = \sum_{\ell,m} f_{\ell m} Y_{\ell m}\p{\v v},
-\end{align}
-where $f_{\ell m}$ are complex coefficients, find a set of points $V=\set{\v v}$ and associated coefficients $\tilde C_{\ell m}\p{\v v}$ with which we can reconstruct $f$, or equivalently its coefficients $f_{\ell m}$, from knowledge of the function's value $f\p{\v v}$ at all points $\v v\in V$; that is
-\begin{align}
-  f_{\ell m} = \sum_{\v v} \tilde C_{\ell m}\p{\v v} f\p{\v v}
-  = \sum_{\v v,k,n} \tilde C_{\ell m}\p{\v v} Y_{kn}\p{\v v} f_{kn}.
+  f\p{\bm v} = \sum_{\ell,m} f_{\ell m} Y_{\ell m}\p{\bm v},
+\end{align}
+where $f_{\ell m}$ are complex coefficients, find a set of points $V=\set{\bm v}$ and associated coefficients $\tilde C_{\ell m}\p{\bm v}$ with which we can reconstruct $f$, or equivalently its coefficients $f_{\ell m}$, from knowledge of the function's value $f\p{\bm v}$ at all points $\bm v\in V$; that is
+\begin{align}
+  f_{\ell m} = \sum_{\bm v} \tilde C_{\ell m}\p{\bm v} f\p{\bm v}
+  = \sum_{\bm v,k,n} \tilde C_{\ell m}\p{\bm v} Y_{kn}\p{\bm v} f_{kn}.
 \end{align}
@@
 \begin{align}
-  \sum_{\v v} \tilde C_{\ell m}\p{\v v} Y_{kn}\p{\v v}
+  \sum_{\bm v} \tilde C_{\ell m}\p{\bm v} Y_{kn}\p{\bm v}
   = \delta_{\ell k} \delta_{mn},
@@
 We will refer to Eq.~\eqref{eq:full_recovery} as the {\it full recovery problem}, and Eq.~\eqref{eq:tomo_recovery} as the the {\it reduced recovery problem}.
-Due to the fact that $D^\ell_{0,m}\p{\v v} = \sqrt{\frac{4\pi}{2\ell+1}}\, Y_{\ell m}\p{\v v}$, any solution to the full recovery problem automatically solves the reduced recovery problem by setting $C_{\ell m}\p{\v v} = \sqrt{\frac{2\ell+1}{4\pi}}\, \tilde C_{\ell m}\p{\v v}^*$.
+Due to the fact that $D^\ell_{0,m}\p{\bm v} = \sqrt{\frac{4\pi}{2\ell+1}}\, Y_{\ell m}\p{\bm v}$, any solution to the full recovery problem automatically solves the reduced recovery problem by setting $C_{\ell m}\p{\bm v} = \sqrt{\frac{2\ell+1}{4\pi}}\, \tilde C_{\ell m}\p{\bm v}^*$.
 In principle, this mapping allows us to import a host of existing signal recovery algorithms \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality} for the task of spin qudit tomography.
@@
 one might expect that $2d-1$ suitably chosen measurement axes would always suffice to perform full tomography of a $d$-level spin qudit.
-In the language of classical signal recovery, we would expect there to exist a set of $2\ell_{\t{max}}+1$ points on the sphere that enable the recovery of any function with fixed degree $\ell\le\ell_{\t{max}}$, given prior knowledge of $\ell$ \footnote{We could further conjecture that if {\it any} set of points $V$ enables the recovery of degree-$\ell_{\t{max}}$ functions on the sphere, then $V$ enables the recovery degree-$\ell$ functions with $\ell\le\ell_{\t{max}}$.
+In the language of classical signal recovery, we would expect there to exist a set of $2\ell_{\t{max}}+1$ points on the sphere that enable the recovery of any function with fixed degree $\ell\le\ell_{\t{max}}$, given prior knowledge of $\ell$\footnote{We could further conjecture that if {\it any} set of points $V$ enables the recovery of degree-$\ell_{\t{max}}$ functions on the sphere, then $V$ enables the recovery degree-$\ell$ functions with $\ell\le\ell_{\t{max}}$.
   However, this condition is stronger than necessary to guarantee full recovery of a $d$-level spin qudit state with $2d-1$ measurement axes, which only requires the existence of {\it some} such set of points.}.
@@
 
-A set of measurement axes $V=\set{\v v}$ nominally induces a set of projectors $\set{\Pi_{\v v\mu}}$ that will be measured in an experiment.
-By a simple change of basis, measuring these projectors is equivalent to measuring the polarization operators $\set{T_{\v v\ell,0}}$.
-Flattening each $d\times d$ matrix $T_{\v v\ell,0}$ into the $d^2$-component column vector $\oket{T_{\v v\ell,0}}$, we construct the {\it measurement matrix}
-\begin{align}
-  M_V \equiv \sum_{\v v,\ell} \ket{\v v\ell} \obra{T_{\v v\ell,0}}.
+A set of measurement axes $V=\set{\bm v}$ nominally induces a set of projectors $\set{\Pi_{\bm v\mu}}$ that will be measured in an experiment.
+By a simple change of basis, measuring these projectors is equivalent to measuring the polarization operators $\set{T_{\bm v\ell,0}}$.
+Flattening each $d\times d$ matrix $T_{\bm v\ell,0}$ into the $d^2$-component column vector $\oket{T_{\bm v\ell,0}}$, we construct the {\it measurement matrix}
+\begin{align}
+  M_V \equiv \sum_{\bm v,\ell} \ket{\bm v\ell} \obra{T_{\bm v\ell,0}}.
   \label{eq:meas_mat}
 \end{align}
-Here $\v v$ and $\ell$ label a row of $M_V$, or equivalently label a standard basis vector $\ket{\v v\ell}$ of a $\p{\abs{V}\times d}$-dimensional vector space, and $\obra{T_{\v v\ell,0}}$ is the conjugate transpose of $\oket{T_{\v v\ell,0}}$.
-A necessary and sufficient condition for $V$ to allow for full state tomography is that the measured polarization operators in $\set{T_{\v v\ell,0}}$, or equivalently the columns of $M_V$, span the entire ($d^2$-dimensional) space of operators on a $d$-level spin qudit.
+Here $\bm v$ and $\ell$ label a row of $M_V$, or equivalently label a standard basis vector $\ket{\bm v\ell}$ of a $\p{\abs{V}\times d}$-dimensional vector space, and $\obra{T_{\bm v\ell,0}}$ is the conjugate transpose of $\oket{T_{\bm v\ell,0}}$.
+A necessary and sufficient condition for $V$ to allow for full state tomography is that the measured polarization operators in $\set{T_{\bm v\ell,0}}$, or equivalently the columns of $M_V$, span the entire ($d^2$-dimensional) space of operators on a $d$-level spin qudit.
 In this case $M_V$ must be full rank, with $d^2$ nonzero singular values.
-Indexing these singular values $M^V_k$ and the corresponding (normalized) left singular vectors $\v x^V_k\equiv \sum_j x^V_{kj} \ket{j}$ by an integer $k\in\set{1,2,\cdots,d^2}$, we can construct the orthonormal qudit operators
+Indexing these singular values $M^V_k$ and the corresponding (normalized) left singular vectors $\bm x^V_k\equiv \sum_j x^V_{kj} \ket{j}$ by an integer $k\in\set{1,2,\cdots,d^2}$, we can construct the orthonormal qudit operators
 \begin{align}
@@
 \end{align}
-where for shorthand we use a combined index $j=\p{\v v,\ell}$ to specify both a measurement axis $\v v$ and a degree $\ell$, which identify the polarization operator $T_j\equiv T_{\v v\ell,0}$.
+where for shorthand we use a combined index $j=\p{\bm v,\ell}$ to specify both a measurement axis $\bm v$ and a degree $\ell$, which identify the polarization operator $T_j\equiv T_{\bm v\ell,0}$.
 These operators allow us to expand any state $\rho$ of a $d$-level spin qudit in the form
@@
 \begin{align}
-  M_{V\ell} \equiv \sum_{\v v,m} \ket{\v v}
-  \obk{T_{\v v\ell,0}|T_{\ell m}} \bra{m}
-  = \sum_{\v v,m} D^\ell_{0,m}\p{\v v} \op{\v v}{m}.
+  M_{V\ell} \equiv \sum_{\bm v,m} \ket{\bm v}
+  \obk{T_{\bm v\ell,0}|T_{\ell m}} \bra{m}
+  = \sum_{\bm v,m} D^\ell_{0,m}\p{\bm v} \op{\bm v}{m}.
   \label{eq:block}
@@
 However, this assumption breaks down when measurement error is limited by fundamental quantum (shot) noise.
-We relax the assumption of Eq.~\eqref{eq:error_assumption} in Appendix \ref{sec:bound}, where we instead assume that $\tilde\rho_V$ is built from $n$ independent measurements of spin projection along every axis $\v v\in V$, with shot noise the dominant source of error.
-In this case, the constraints that $\tr\p\rho=1$ and $\bk{\Pi_{\v v\mu}}_\rho\ge0$ allow us to bound the mean squared distance between $\tilde\rho_V$ and $\rho$ as
+We relax the assumption of Eq.~\eqref{eq:error_assumption} in Appendix \ref{sec:bound}, where we instead assume that $\tilde\rho_V$ is built from $n$ independent measurements of spin projection along every axis $\bm v\in V$, with shot noise the dominant source of error.
+In this case, the constraints that $\tr\p\rho=1$ and $\bk{\Pi_{\bm v\mu}}_\rho\ge0$ allow us to bound the mean squared distance between $\tilde\rho_V$ and $\rho$ as
 \begin{align}
@@
 
-Though straightforward to compute, the bound in Eq.~\eqref{eq:bound} is not tight, as it is acquired by bounding the statistical error in the empirical estimate $\tilde T_{\v v\ell,0}$ of $\bk{T_{\v v\ell,0}}_\rho$ by $\bbk{\epsilon_{\v v\ell,0}^2}\le\Gamma_\ell^2$.
-The individual bounds on $\bbk{\epsilon_{\v v\ell,0}^2}$ for each axis $\v v$ and degree $\ell$ are tight, but these bounds cannot all be achieved simultaneously.
+Though straightforward to compute, the bound in Eq.~\eqref{eq:bound} is not tight, as it is acquired by bounding the statistical error in the empirical estimate $\tilde T_{\bm v\ell,0}$ of $\bk{T_{\bm v\ell,0}}_\rho$ by $\bbk{\epsilon_{\bm v\ell,0}^2}\le\Gamma_\ell^2$.
+The individual bounds on $\bbk{\epsilon_{\bm v\ell,0}^2}$ for each axis $\bm v$ and degree $\ell$ are tight, but these bounds cannot all be achieved simultaneously.
 There is therefore still room for improvement on the bound in Eq.~\eqref{eq:bound} by maximizing $\E_V$ over the set of all qudit states $\rho$.
@@
 \begin{align}
-  \E_V\p{\rho}^2 = \sum_{\v v,\v w,\ell}
-  \bk{\v v|\p{M_{V\ell}^{-1}}^\dag M_{V\ell}^{-1}|\v w}
-  \bbk{\epsilon_{\v v\ell}\epsilon_{\v w\ell}},
+  \E_V\p{\rho}^2 = \sum_{\bm v,\bm w,\ell}
+  \bk{\bm v|\p{M_{V\ell}^{-1}}^\dag M_{V\ell}^{-1}|\bm w}
+  \bbk{\epsilon_{\bm v\ell}\epsilon_{\bm w\ell}},
   \label{eq:error_cls}
 \end{align}
-where $\bbk{\epsilon_{\v v\ell}\epsilon_{\v w\ell}}$ is the covariance between errors $\epsilon_{\v v\ell}$ in the empirical estimates $\tilde T_{\v v\ell,0}$ of $\bk{T_{\v v\ell,0}}_\rho$.
-These covariances are generally determined by the sources of measurement error in any given experiment, but will typically satisfy $\bbk{\epsilon_{\v v\ell}\epsilon_{\v w\ell}} = \delta_{\v v\v w}\bbk{\epsilon_{\v v\ell}^2}$.
+where $\bbk{\epsilon_{\bm v\ell}\epsilon_{\bm w\ell}}$ is the covariance between errors $\epsilon_{\bm v\ell}$ in the empirical estimates $\tilde T_{\bm v\ell,0}$ of $\bk{T_{\bm v\ell,0}}_\rho$.
+These covariances are generally determined by the sources of measurement error in any given experiment, but will typically satisfy $\bbk{\epsilon_{\bm v\ell}\epsilon_{\bm w\ell}} = \delta_{\bm v\bm w}\bbk{\epsilon_{\bm v\ell}^2}$.
 If measurement error is limited by shot noise, then (see Appendix \ref{sec:exact})
@@
 \begin{enumerate}
-\item Select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and use any standard minimization algorithm to optimize the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) by minimizing the quantum error scale $\epsilon_V$ in Eq.~\eqref{eq:bound}.
+\item Select a random set of measurement axes $V$ by uniformly sampling points on the sphere\footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and use any standard minimization algorithm to optimize the $2\abs{V}$ parameters in $V$ (two angles for each point $\bm v\in V$) by minimizing the quantum error scale $\epsilon_V$ in Eq.~\eqref{eq:bound}.
   If $\abs{V}$ is too large for such optimization, you can simply generate many sets of random measurement axes, and then choose the set with the smallest quantum error scale $\epsilon_V$.
@@
   Save all measurement matrix data associated with the final measurement axes $V$ for later use.
-\item For each axis $\v v\in V$, make $n$ measurements of spin projection, and set $\tilde\Pi_{\v v\mu}\approx\bk{\Pi_{\v v\mu}}_\rho$ to the fraction of times in which the measurement outcome was $\mu$.
-\item Use the the estimates $\tilde\Pi_{\v v\mu}$ of $\bk{\Pi_{\v v\mu}}_\rho$ to compute the estimates
-\begin{align}
-  \tilde T_{\v v\ell,0}
-  \equiv \sum_\mu \bk{\mu|T_{\ell,0}|\mu} \tilde\Pi_{\v v\mu}
-\end{align}
-of $\bk{T_{\v v\ell,0}}_\rho$, where the matrix elements of $T_{\ell,0}$ are provided in Eq.~\eqref{eq:trans_op}.
-\item Denoting the nonzero singular values of $M_{V\ell}$ by $M^V_{\ell k}$ and the corresponding left singular vectors by $\v x^V_{\ell k} = \sum_{\v v} x^V_{\ell k \v v} \ket{\v v}$, compute the  operators and coefficients
-  \begin{align}
-    Q_{\ell k} &\equiv \f1{M^V_{\ell k}} \sum_{\v v}
-    \p{x^V_{\ell k\v v}}^* T_{\v v\ell,0}, \\
-    \tilde\rho^V_{\ell k} &\equiv \f1{M^V_{\ell k}} \sum_{\v v}
-    x^V_{\ell k\v v} \, \tilde T_{\v v\ell,0},
+\item For each axis $\bm v\in V$, make $n$ measurements of spin projection, and set $\tilde\Pi_{\bm v\mu}\approx\bk{\Pi_{\bm v\mu}}_\rho$ to the fraction of times in which the measurement outcome was $\mu$.
+\item Use the the estimates $\tilde\Pi_{\bm v\mu}$ of $\bk{\Pi_{\bm v\mu}}_\rho$ to compute the estimates
+\begin{align}
+  \tilde T_{\bm v\ell,0}
+  \equiv \sum_\mu \bk{\mu|T_{\ell,0}|\mu} \tilde\Pi_{\bm v\mu}
+\end{align}
+of $\bk{T_{\bm v\ell,0}}_\rho$, where the matrix elements of $T_{\ell,0}$ are provided in Eq.~\eqref{eq:trans_op}.
+\item Denoting the nonzero singular values of $M_{V\ell}$ by $M^V_{\ell k}$ and the corresponding left singular vectors by $\bm x^V_{\ell k} = \sum_{\bm v} x^V_{\ell k \bm v} \ket{\bm v}$, compute the  operators and coefficients
+  \begin{align}
+    Q_{\ell k} &\equiv \f1{M^V_{\ell k}} \sum_{\bm v}
+    \p{x^V_{\ell k\bm v}}^* T_{\bm v\ell,0}, \\
+    \tilde\rho^V_{\ell k} &\equiv \f1{M^V_{\ell k}} \sum_{\bm v}
+    x^V_{\ell k\bm v} \, \tilde T_{\bm v\ell,0},
   \end{align}
@@
 
-In principle, any information about $\rho$ obtained from prior knowledge or preliminary measurement data can be used to construct tailored or adaptive measurement protocols \cite{pereira2018adaptive} that are more efficient in terms of the number of measurements required to estimate $\rho$ to a fixed precision.
+In principle, any information about $\rho$ obtained from prior knowledge or preliminary measurement data can be used to construct tailored or adaptive measurement protocols \cite{huszar2012adaptive, ferrie2014selfguided, granade2016practical, pereira2018adaptive} that are more efficient in terms of the number of measurements required to estimate $\rho$ to a fixed precision.
 We leave the development of tailored and adaptive measurement protocols to future work.
@@
 Increasing the number of measurement axes generally decreases the quantum error scale $\epsilon_V$, but comes at the cost of having to estimate more observables.
-At a fixed total number of measurements, increasing $\abs{V}$ reduces the number of measurements $n$ devoted to each axis $\v v\in V$.
+At a fixed total number of measurements, increasing $\abs{V}$ reduces the number of measurements $n$ devoted to each axis $\bm v\in V$.
 This trade-off begs the question: how should one choose the number of measurement axes, $\abs{V}$?
@@
 
-\bibliography{\jobname.bib}
-
-\onecolumngrid
+\printbibliography
+\onecolumn
+\newpage
 \appendix
@@
   &&
-  \v S \equiv \p{S_\x,S_\y,S_\z}.
-\end{align}
-The spin vector $\v S$ generates rotations of a spin-$s$ system in 3D space.
-Specifically, the operator $e^{-\i\theta\v S\cdot\uv n}$ rotates a spin-$s$ system by an angle $\theta$ about the unit vector $\uv n$.
+  \bm S \equiv \p{S_\x,S_\y,S_\z}.
+\end{align}
+The spin vector $\bm S$ generates rotations of a spin-$s$ system in 3D space.
+Specifically, the operator $e^{-\i\theta\bm S\cdot\uv n}$ rotates a spin-$s$ system by an angle $\theta$ about the unit vector $\uv n$.
 
@@
 \end{align}
-which implies that $T_{\ell m}$ is a spherical tensor operator, whose degree degree $\ell$ is preserved under rotations generated by $\v S$.
+which implies that $T_{\ell m}$ is a spherical tensor operator, whose degree degree $\ell$ is preserved under rotations generated by $\bm S$.
 Moreover, by comparing Eqs.~\eqref{eq:spin_ops} and \eqref{eq:spin_trans} we see that the polarization operators $T_{\ell m}$ transform identically to spin-$\ell$ particles under the adjoint action of the spin operators $S_\z$ and $S_\pm$.
-For any triplet of angles $\v\omega=\p{\alpha,\beta,\gamma}$, we can therefore define the rotation operator
-\begin{align}
-  R\p{\v\omega} \equiv
+For any triplet of angles $\bm\omega=\p{\alpha,\beta,\gamma}$, we can therefore define the rotation operator
+\begin{align}
+  R\p{\bm\omega} \equiv
   e^{-\i\alpha S_\z} e^{-\i\beta S_\y} e^{-\i\gamma S_\z},
@@
 \begin{align}
-  T_{\v\omega\ell m} \equiv
-  R\p{\v\omega} T_{\ell m} R\p{\v\omega}^\dag
-  = \sum_{n=-\ell}^\ell D_{mn}^\ell\p{\bar{\v\omega}}^* T_{\ell n},
+  T_{\bm\omega\ell m} \equiv
+  R\p{\bm\omega} T_{\ell m} R\p{\bm\omega}^\dag
+  = \sum_{n=-\ell}^\ell D_{mn}^\ell\p{\bar{\bm\omega}}^* T_{\ell n},
   \label{eq:trans_rot_apndx}
 \end{align}
-where $\bar{\v\omega}=\p{\gamma,\beta,\alpha}$ is the reversal of $\v\omega$, and
-\begin{align}
-  D_{mn}^\ell\p{\bar{\v\omega}}
-  \equiv \bk{\ell m|R\p{\bar{\v\omega}}|\ell n}
-  = \obk{T_{\ell n} | R\p{\v\omega}
-    \otimes R\p{\v\omega}^* |T_{\ell m}}^*
-  = \obk{T_{\ell m} | R\p{-\bar{\v\omega}}
-    \otimes R\p{-\bar{\v\omega}}^* |T_{\ell n}}
-\end{align}
-are matrix elements of the rotation operator $R\p{\v\omega}$ for spin-$\ell$ particles.
-
-For any angle doublet $\v v=\p{\alpha,\beta}$, we define $R\p{\v v}\equiv R\p{\alpha,\beta,0}$ and $D^\ell_{mn}\p{\v v}=D^\ell_{mn}\p{0,\beta,\alpha}$ for shorthand.
+where $\bar{\bm\omega}=\p{\gamma,\beta,\alpha}$ is the reversal of $\bm\omega$, and
+\begin{align}
+  D_{mn}^\ell\p{\bar{\bm\omega}}
+  \equiv \bk{\ell m|R\p{\bar{\bm\omega}}|\ell n}
+  = \obk{T_{\ell n} | R\p{\bm\omega}
+    \otimes R\p{\bm\omega}^* |T_{\ell m}}^*
+  = \obk{T_{\ell m} | R\p{-\bar{\bm\omega}}
+    \otimes R\p{-\bar{\bm\omega}}^* |T_{\ell n}}
+\end{align}
+are matrix elements of the rotation operator $R\p{\bm\omega}$ for spin-$\ell$ particles.
+
+For any angle doublet $\bm v=\p{\alpha,\beta}$, we define $R\p{\bm v}\equiv R\p{\alpha,\beta,0}$ and $D^\ell_{mn}\p{\bm v}=D^\ell_{mn}\p{0,\beta,\alpha}$ for shorthand.
 The transformation rules in Eq.~\eqref{eq:trans_rot_apndx} imply that we can expand the phase-space representation of $T_{\ell m}$ as
 \begin{align}
-  T_{\ell m}^{\t{PS}}\p{\v v}
-  \equiv \bk{s_{\v v} | T_{\ell m} | s_{\v v}}
-  = \bk{s | R\p{\v v}^\dag T_{\ell m} R\p{\v v} | s}
-  = D^\ell_{0,m}\p{\v v} \bk{s | T_{\ell,0} | s},
+  T_{\ell m}^{\t{PS}}\p{\bm v}
+  \equiv \bk{s_{\bm v} | T_{\ell m} | s_{\bm v}}
+  = \bk{s | R\p{\bm v}^\dag T_{\ell m} R\p{\bm v} | s}
+  = D^\ell_{0,m}\p{\bm v} \bk{s | T_{\ell,0} | s},
 \end{align}
@@
 \begin{align}
-  D^\ell_{0,m}\p{\v v} = \sqrt{\f{4\pi}{2\ell+1}} \, Y_{\ell m}\p{\v v},
+  D^\ell_{0,m}\p{\bm v} = \sqrt{\f{4\pi}{2\ell+1}} \, Y_{\ell m}\p{\bm v},
 \end{align}
@@
 \begin{align}
-  T_{\ell m}^{\t{PS}}\p{\v v}
+  T_{\ell m}^{\t{PS}}\p{\bm v}
   = \sqrt{\f{4\pi}{2s+\ell+1}
     \p{\f{\p{2s}!}{\p{2s+\ell}!}} \p{\f{\p{2s}!}{\p{2s-\ell}!}}} \,
-  Y_{\ell m}\p{\v v}.
+  Y_{\ell m}\p{\bm v}.
 \end{align}
@@
 
-To this end, we fix a particular set of measurement axes $V$, and consider performing $n$ measurements of spin projection along every axis $\v v\in V$, for a total of $N=\abs{V}\times n$ measurements.
+To this end, we fix a particular set of measurement axes $V$, and consider performing $n$ measurements of spin projection along every axis $\bm v\in V$, for a total of $N=\abs{V}\times n$ measurements.
 Such a procedure is equivalent to making $N$ local measurements of the $N$-fold product state $\rho^{\otimes N}$.
-For convenience, we index the tensor factors of $\rho^{\otimes N}$ by the integers $\p{i,j}$, with $i\in\set{1,2,\cdots,\abs{V}}$ specifying a measurement axis $\v v_i\in V$, and $j\in\set{1,2,\cdots,n}$ specifying the copy of $\rho$ prepared for the $j$-th measurement spin projection along a particular axis.
-We then define the projectors $\Pi_{i\mu}\equiv\op{\mu_{\v v_i}}$ onto states $\ket{\mu_{\v v_i}}$ with definite spin projection $\mu$ along axis $\v v_i\in V$, and define $\Pi_{i\mu}^j$ to be an $N$-qudit operator with $\Pi_{i\mu}$ on the $\p{i,j}$-th tensor factor and the identity elsewhere.
+For convenience, we index the tensor factors of $\rho^{\otimes N}$ by the integers $\p{i,j}$, with $i\in\set{1,2,\cdots,\abs{V}}$ specifying a measurement axis $\bm v_i\in V$, and $j\in\set{1,2,\cdots,n}$ specifying the copy of $\rho$ prepared for the $j$-th measurement spin projection along a particular axis.
+We then define the projectors $\Pi_{i\mu}\equiv\op{\mu_{\bm v_i}}$ onto states $\ket{\mu_{\bm v_i}}$ with definite spin projection $\mu$ along axis $\bm v_i\in V$, and define $\Pi_{i\mu}^j$ to be an $N$-qudit operator with $\Pi_{i\mu}$ on the $\p{i,j}$-th tensor factor and the identity elsewhere.
 We denote the experimental outcome of measuring $\Pi_{i\mu}$ in the $\p{i,j}$-th copy of $\rho$ by $\tilde\Pi_{i\mu}^j\in\set{0,1}$.
 In other words, $\tilde\Pi_{i\mu}^j$ is the ``single-shot estimate'' of $\Pi_{i\mu}$, with $\tilde\Pi_{i\mu}^j=1$ if outcome $\mu$ was observed on the $\p{i,j}$-th experimental trial, and $\tilde\Pi_{i\mu}^j=0$ otherwise.
-An empirical estimate of the expectation value $\bk{\Pi_{i\mu}}_\rho$ is provided by the fraction of times that outcome $\mu$ was observed when measuring spin projection along axis $\v v_i$, that is
+An empirical estimate of the expectation value $\bk{\Pi_{i\mu}}_\rho$ is provided by the fraction of times that outcome $\mu$ was observed when measuring spin projection along axis $\bm v_i$, that is
 \begin{align}
@@
 \end{align}
-is the average of $\Pi_{i\mu}$ applied to all copies of $\rho$ for which spin projection is measured along the axis $\v v_i$.
+is the average of $\Pi_{i\mu}$ applied to all copies of $\rho$ for which spin projection is measured along the axis $\bm v_i$.
 Eq.~\eqref{eq:proj_estimate} implies that
@@
 
-Rather than the statistical errors $\epsilon_{i\mu} \equiv \tilde\Pi_{i\mu} - \bk{\Pi_{i\mu}}_\rho$ in the estimates $\tilde\Pi_{i\mu}$ of the projectors $\Pi_{i\mu}$, we now consider the statistical errors $\epsilon_{i\ell} \equiv \tilde T_{i\ell} - \bk{T_{i\ell}}_\rho$ in the estimates $\tilde T_{i\ell}$ of the polarization operators $T_{i\ell} \equiv T_{\v v_i\ell,0}$.
+Rather than the statistical errors $\epsilon_{i\mu} \equiv \tilde\Pi_{i\mu} - \bk{\Pi_{i\mu}}_\rho$ in the estimates $\tilde\Pi_{i\mu}$ of the projectors $\Pi_{i\mu}$, we now consider the statistical errors $\epsilon_{i\ell} \equiv \tilde T_{i\ell} - \bk{T_{i\ell}}_\rho$ in the estimates $\tilde T_{i\ell}$ of the polarization operators $T_{i\ell} \equiv T_{\bm v_i\ell,0}$.
 We can expand the polarization operators $T_{i\ell}$ as a sum over projectors $\Pi_{i\mu}$ as
@@
 We now revisit the derivation of reconstruction error in Section \ref{sec:error} to make use of the bounds on variances $\bbk{\epsilon_{i\ell}^2}$.
-To recap, for a set of measurement axes $V=\set{\v v}$ and degrees $\ell\in\set{0,1,\cdots,d-1}$ we construct the measurement matrix
-\begin{align}
-  M_V \equiv \sum_{\v v,\ell} \ket{\v v\ell} \obra{T_{\v v\ell,0}},
+To recap, for a set of measurement axes $V=\set{\bm v}$ and degrees $\ell\in\set{0,1,\cdots,d-1}$ we construct the measurement matrix
+\begin{align}
+  M_V \equiv \sum_{\bm v,\ell} \ket{\bm v\ell} \obra{T_{\bm v\ell,0}},
 \end{align}
@@
   &&
-  M_{V\ell} = \sum_{m,\v v} D^\ell_{m,0}\p{\v v} \op{\v v}{m},
-\end{align}
-where $D^\ell_{mn}\p{\v v}\equiv\bk{\ell m|R\p{\v v}|\ell n}$ is a (Wigner) rotation matrix element for a spin-$\ell$ particle.
-The block-diagonal structure of $M_V$ allows us to index its singular values $M^V_{\ell m}$ and corresponding (normalized) left singular vectors $\v x^V_{\ell m} = \sum_i x^V_{\ell m i} \ket{\v v_i}$ by the indices $\p{\ell,m}$, where the integer $\abs{m}\le\ell$.
+  M_{V\ell} = \sum_{m,\bm v} D^\ell_{m,0}\p{\bm v} \op{\bm v}{m},
+\end{align}
+where $D^\ell_{mn}\p{\bm v}\equiv\bk{\ell m|R\p{\bm v}|\ell n}$ is a (Wigner) rotation matrix element for a spin-$\ell$ particle.
+The block-diagonal structure of $M_V$ allows us to index its singular values $M^V_{\ell m}$ and corresponding (normalized) left singular vectors $\bm x^V_{\ell m} = \sum_i x^V_{\ell m i} \ket{\bm v_i}$ by the indices $\p{\ell,m}$, where the integer $\abs{m}\le\ell$.
 These singular vectors and values define the orthonormal operators
@@
 \end{align}
-where $i\in\set{1,2,\cdots,\abs{V}}$ indexes an axis $\v v_i\in V$, with $T_{i\ell} \equiv T_{\v v_i\ell}$.
+where $i\in\set{1,2,\cdots,\abs{V}}$ indexes an axis $\bm v_i\in V$, with $T_{i\ell} \equiv T_{\bm v_i\ell}$.
 The state $\rho$ can be expanded in the basis of these operators as
@@
   = \sum_{\ell,m,i,i'} \p{q^V_{\ell mi}}^* q^V_{\ell mi'} \bbk{\epsilon_{i\ell}\epsilon_{i'\ell}}
-  = \f1n \sum_{\ell,i} \abs{\tilde{\v q}_{\ell i}}^2 \cov_\rho\p{T_{i\ell},T_{i\ell}},
-  &&
-  \abs{\tilde{\v q}_{\ell i}}^2 = \sum_m \abs{q_{\ell mi}}^2,
+  = \f1n \sum_{\ell,i} \abs{\tilde{\bm q}_{\ell i}}^2 \cov_\rho\p{T_{i\ell},T_{i\ell}},
+  &&
+  \abs{\tilde{\bm q}_{\ell i}}^2 = \sum_m \abs{q_{\ell mi}}^2,
   \label{eq:error_SM_start}
 \end{align}
-where $\tilde{\v q}_{\ell i} = \sum_m \p{q_{\ell mi}}^* \ket{m}$, and we used the fact that $\bbk{\epsilon_{i\ell}\epsilon_{i'\ell}} = \delta_{ii'}\times\cov_\rho\p{T_{i\ell},T_{i\ell}}/n$.
-Identifying the singular value decomposition $M_{V\ell} = U_{V\ell} \Sigma_{V\ell} W_{V\ell}^\dag$, we then we observe that $\tilde{\v q}_{\ell i} = \Sigma_{V\ell}^{-1} U_{V\ell}^\dag \ket{\v v_i}$, which allows us to simplify
-\begin{align}
-  \abs{\tilde{\v q}_{\ell i}}^2
-  = \bk{\v v_i | U_{V\ell} \Sigma_{V\ell}^{-2} U_{V\ell}^\dag | \v v_i}
-  = \bk{\v v_i| \p{M_{V\ell}^{-1}}^\dag M_{V\ell}^{-1} | \v v_i}.
+where $\tilde{\bm q}_{\ell i} = \sum_m \p{q_{\ell mi}}^* \ket{m}$, and we used the fact that $\bbk{\epsilon_{i\ell}\epsilon_{i'\ell}} = \delta_{ii'}\times\cov_\rho\p{T_{i\ell},T_{i\ell}}/n$.
+Identifying the singular value decomposition $M_{V\ell} = U_{V\ell} \Sigma_{V\ell} W_{V\ell}^\dag$, we then we observe that $\tilde{\bm q}_{\ell i} = \Sigma_{V\ell}^{-1} U_{V\ell}^\dag \ket{\bm v_i}$, which allows us to simplify
+\begin{align}
+  \abs{\tilde{\bm q}_{\ell i}}^2
+  = \bk{\bm v_i | U_{V\ell} \Sigma_{V\ell}^{-2} U_{V\ell}^\dag | \bm v_i}
+  = \bk{\bm v_i| \p{M_{V\ell}^{-1}}^\dag M_{V\ell}^{-1} | \bm v_i}.
 \end{align}
@@
   = \cov_\rho\p{T_{i\ell}^\dag,T_{i\ell}}
-  = \sum_{m,m'} D^\ell_{0,m}\p{\v v_i} D^\ell_{0,m'}\p{\v v_i}^*
+  = \sum_{m,m'} D^\ell_{0,m}\p{\bm v_i} D^\ell_{0,m'}\p{\bm v_i}^*
   \cov_\rho\p{T_{\ell m}^\dag, T_{\ell m'}},
@@
   = \f1n \sum_{\ell,i,m,m'}
-  D^\ell_{0,m'}\p{\v v_i}^* \abs{\tilde{\v q}_{\ell i}}^2
-  D^\ell_{0,m}\p{\v v_i}
+  D^\ell_{0,m'}\p{\bm v_i}^* \abs{\tilde{\bm q}_{\ell i}}^2
+  D^\ell_{0,m}\p{\bm v_i}
   \cov_\rho\p{T_{\ell m}^\dag, T_{\ell m'}}.
@@
 we can use the fact that $2\ell_2$ is always even (because $\ell_2$ is always an integer) to expand
-\begin{align}
+\begin{multline}
   f_{\ell_1 m_1;\ell_2 m_2}^{LM}
-  = \p{-1}^M \sqrt{\p{2L+1}\p{2\ell_1+1}\p{2\ell_2+1}}
-  \sum_{\mu,\nu,\rho} \p{-1}^{3s-\mu-\nu-\rho}
+  = \p{-1}^M \sqrt{\p{2L+1}\p{2\ell_1+1}\p{2\ell_2+1}} \\
+  \times \sum_{\mu,\nu,\rho} \p{-1}^{3s-\mu-\nu-\rho}
   \begin{pmatrix}
@@
   \end{pmatrix}.
-\end{align}
+\end{multline}
 This sum can be simplified by the introduction of Wigner 6-$j$ symbols, giving us

### 2020-12-11 sun_phases/spin_qudit_tomo.tex
final changes for first arxiv version of spin tomography paper

@@
 \title{Spin qudit tomography}
-\author{Michael A.~Perlin}
+\author[]{Michael A.~Perlin}
 \email{mika.perlin@gmail.com}
 \orcid{https://orcid.org/0000-0002-9316-1596}
-\author{Diego Barberena}
-\author{Ana Maria Rey}
+\author[]{Diego Barberena}
+\author[]{Ana Maria Rey}
 \orcid{http://orcid.org/0000-0001-7176-9413}
-\affiliation{JILA, National Institute of Standards and Technology and University of Colorado, 440 UCB, Boulder, Colorado 80309, USA}
-\affiliation{Center for Theory of Quantum Matter, University of Colorado, Boulder, CO, 80309, USA}
-\date{10 December 2020}
+\affiliation[1]{JILA, National Institute of Standards and Technology and University of Colorado, 440 UCB, Boulder, Colorado 80309, USA}
+\affiliation[2]{Center for Theory of Quantum Matter, University of Colorado, Boulder, CO, 80309, USA}
 
@@
 In this work, we provide an efficient framework for performing spin qudit tomography using only measurements of spin projection onto different spatial axes.
-We first introduce, in Section \ref{sec:polarization_ops}, a set of qudit operators that are a quantum analogue of the spherical harmonics, and which play a central role in our work.
+We first introduce, in Section \ref{sec:polarization_ops}, a set of qudit operators that are closely related to the spherical harmonics, and which play a central role in our work.
 We then map the quantum problem of spin qudit tomography onto the well-studied classical problem of signal recovery on the sphere in Section \ref{sec:signal_recovery}, thereby importing a host of existing literature and mathematical machinery that has been developed for this task \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
@@
 
-We begin by introducing a set of qudit operators that are a quantum analogue of the spherical harmonics (in a sense that will be clarified below), and which play a central role in our work.
+We begin by introducing a set of qudit operators that are closely related to the spherical harmonics (in a sense that will be clarified below), and which play a central role in our work.
 Consider a $d$-state spin qudit with total spin $s\equiv\frac{d-1}{2}$.
@@
 The polarization operators can be interpreted in terms of an absorption process, whereby $T_{\ell m}\ket\psi$ is (up to normalization) the state obtained after a spin-$s$ state $\ket\psi$ absorbs a particle with total spin $\ell$ and spin projection $m$ onto a fixed quantization axis.
-In analogy with the complex spherical harmonics $Y_{\ell m}$, we will refer to $\ell$ as the {\it degree} and $m$ as the {\it order} of $T_{\ell m}$.
+Similarly to the complex spherical harmonics $Y_{\ell m}$, we will refer to $\ell$ as the {\it degree} and $m$ as the {\it order} of $T_{\ell m}$.
 
@@
 \end{align}
-Expanding the rotated polarization operators $T_{\bm vk,0}$ into a sum of un-rotated polarization operators, according to Eq.~\eqref{eq:trans_rot}, we find that the recovery condition in Eq.~\eqref{eq:state_recon} is satisfied when
+Expanding the rotated polarization operators $T_{\bm vk,0}$ into a sum of un-rotated polarization operators $T_{\ell n}$ according to Eq.~\eqref{eq:trans_rot}, we find that the recovery condition in Eq.~\eqref{eq:state_recon} is satisfied when
 \begin{align}
@@
 If the function $f$ is {\it band-limited} at degree $L$, which is to say that $f_{\ell m}=0$ for all $\ell\ge L$, then the full recovery problem in Eq.~\eqref{eq:full_recovery} is provably solvable with a suitable choice of $\abs{V}=L^2$ points on the sphere \cite{freeden2008spherical, freeden2018spherical}.
-The existence of these solutions to the full recovery problem in turn implies the existence of $d^2$ measurement axes that suffice to reconstruct arbitrary states of $d$-level spin qudit, whose possible states (or rather, their phase-space representations) are band-limited at degree $d$.
+The existence of these solutions to the full recovery problem in turn implies the existence of $d^2$ measurement axes that suffice to reconstruct arbitrary states of $d$-level spin qudit, whose possible states (or rather, phase-space representations) are band-limited at degree $d$.
 Moreover, for any fixed degree $\ell$, finding solutions to the reduced recovery problem in Eq.~\eqref{eq:tomo_recovery} is equivalent to the recovery of a degree-$\ell$ function $f_\ell = \sum_m f_{\ell m} Y_{\ell m}$, which is provably possible with $\abs{V}=2\ell+1$ samples \cite{freeden2008spherical}.
@@
 \end{enumerate*}
-Whereas step \ref{tomo:measure} can involve a host of platform-dependent technical challenges, in the following two sections we discuss the steps to take before and after collecting measurement data.
+Whereas step \ref{tomo:measure} can involve a host of platform-dependent technical challenges, in the following sections we discuss the steps to take before and after collecting measurement data.
 
@@
 \end{align}
-where, using the fact that $T_j=T_j^\dag$,
+where, using the fact that $T_j=T_j^\dag$ (because they are degree $m=0$ polarization operators),
 \begin{align}
@@
 
-Though straightforward to compute, the bound in Eq.~\eqref{eq:bound} is not tight, as it is acquired by bounding the statistical error in the empirical estimate $\tilde T_{\bm v\ell,0}$ of $\bk{T_{\bm v\ell,0}}_\rho$ by $\bbk{\epsilon_{\bm v\ell,0}^2}\le\Gamma_\ell^2$.
-The individual bounds on $\bbk{\epsilon_{\bm v\ell,0}^2}$ for each axis $\bm v$ and degree $\ell$ are tight, but these bounds cannot all be achieved simultaneously.
+Though straightforward to compute, the bound in Eq.~\eqref{eq:bound} is not tight, as it is acquired by bounding the statistical error $\epsilon_{\bm v\ell}$ in the empirical estimate $\tilde T_{\bm v\ell,0}$ of $\bk{T_{\bm v\ell,0}}_\rho$ by $\bbk{\epsilon_{\bm v\ell}^2}\le\Gamma_\ell^2$.
+The individual bounds on $\bbk{\epsilon_{\bm v\ell}^2}$ for each axis $\bm v$ and degree $\ell$ are tight, but these bounds cannot all be achieved simultaneously.
 There is therefore still room for improvement on the bound in Eq.~\eqref{eq:bound} by maximizing $\E_V$ over the set of all qudit states $\rho$.
-We discuss this maximization problem in Appendix \ref{sec:exact}.
+We discuss this maximization problem in Appendix \ref{sec:exact}, but leave its full solution to future work.
 We also note that the (squared) reconstruction error bound in Eq.~\eqref{eq:bound} obeys the ``standard quantum limit'' of $\sim1/n$ scaling in the number of measurements.
@@
   \bk{\bm v|\p{M_{V\ell}^{-1}}^\dag M_{V\ell}^{-1}|\bm w}
-  \bbk{\epsilon_{\bm v\ell}\epsilon_{\bm w\ell}},
+  \bbk{\epsilon_{\bm v\ell}\epsilon_{\bm w\ell}}.
   \label{eq:error_cls}
 \end{align}
-where $\bbk{\epsilon_{\bm v\ell}\epsilon_{\bm w\ell}}$ is the covariance between errors $\epsilon_{\bm v\ell}$ in the empirical estimates $\tilde T_{\bm v\ell,0}$ of $\bk{T_{\bm v\ell,0}}_\rho$.
-These covariances are generally determined by the sources of measurement error in any given experiment, but will typically satisfy $\bbk{\epsilon_{\bm v\ell}\epsilon_{\bm w\ell}} = \delta_{\bm v\bm w}\bbk{\epsilon_{\bm v\ell}^2}$.
+The covariances $\bbk{\epsilon_{\bm v\ell}\epsilon_{\bm w\ell}}$ are generally determined by the sources of measurement error in any given experiment, but will typically satisfy $\bbk{\epsilon_{\bm v\ell}\epsilon_{\bm w\ell}} = \delta_{\bm v\bm w}\bbk{\epsilon_{\bm v\ell}^2}$.
 If measurement error is limited by shot noise, then (see Appendix \ref{sec:exact})
@@
 \end{align}
-where $\stackrel{\t{SNL}}{=}$ indicates equality in the ``shot-noise-limited'' regime; $\ket{\rho_\ell} \equiv \sum_m \rho_{\ell m}\ket{m}$ is a vector of the scalar polarization operator components $\rho_{\ell m}$ of $\rho$, defined in Eq.~\eqref{eq:trans_state}; and the matrix $\N_{V\ell}$ and vector $\ket{\chi_{V\ell}}$ are defined below.
+where $\stackrel{\t{SNL}}{=}$ indicates equality in the ``shot-noise-limited'' regime; $\ket{\rho_\ell} \equiv \sum_m \rho_{\ell m}\ket{m}$ is a vector of the polarization operator components $\rho_{\ell m}$ of $\rho$, defined in Eq.~\eqref{eq:trans_state}; and the matrix $\N_{V\ell}$ and vector $\ket{\chi_{V\ell}}$ are defined below.
 While the true shot-noise-limited error in $\tilde\rho_V$ cannot be known exactly without knowing $\rho$, this error can be estimated a posteriori by $\E_V\p{\rho}\approx\E_V\p{\tilde\rho_V}$.
@@
 which implies that $T_{\ell m}$ is a spherical tensor operator, whose degree degree $\ell$ is preserved under rotations generated by $\bm S$.
-Moreover, by comparing Eqs.~\eqref{eq:spin_ops} and \eqref{eq:spin_trans} we see that the polarization operators $T_{\ell m}$ transform identically to spin-$\ell$ particles under the adjoint action of the spin operators $S_\z$ and $S_\pm$.
+Moreover, by comparing Eqs.~\eqref{eq:spin_ops} and \eqref{eq:spin_trans} we see that the polarization operators $T_{\ell m}$ transform identically to spin-$\ell$ particles under the (adjoint) action of the spin operators $S_\z$ and $S_\pm$.
 For any triplet of angles $\bm\omega=\p{\alpha,\beta,\gamma}$, we can therefore define the rotation operator
@@
 \item there is no a priori guarantee for the value of $\epsilon$, which must be inferred from experimental outcomes, and
-\item the assumption that all errors are uncorrelated is unjustified (and generally false).
+\item the assumption that all measurement errors are uncorrelated is unjustified (and generally false).
 \end{enumerate*}
@@
 For convenience, we index the tensor factors of $\rho^{\otimes N}$ by the integers $\p{i,j}$, with $i\in\set{1,2,\cdots,\abs{V}}$ specifying a measurement axis $\bm v_i\in V$, and $j\in\set{1,2,\cdots,n}$ specifying the copy of $\rho$ prepared for the $j$-th measurement spin projection along a particular axis.
-We then define the projectors $\Pi_{i\mu}\equiv\op{\mu_{\bm v_i}}$ onto states $\ket{\mu_{\bm v_i}}$ with definite spin projection $\mu$ along axis $\bm v_i\in V$, and define $\Pi_{i\mu}^j$ to be an $N$-qudit operator with $\Pi_{i\mu}$ on the $\p{i,j}$-th tensor factor and the identity elsewhere.
+We then define the projectors $\Pi_{i\mu}\equiv\op{\mu_{\bm v_i}}$ onto single-qudit states $\ket{\mu_{\bm v_i}}$ with definite spin projection $\mu$ along axis $\bm v_i\in V$, and define $\Pi_{i\mu}^j$ to be an $N$-qudit operator with $\Pi_{i\mu}$ on the $\p{i,j}$-th tensor factor and the identity elsewhere.
 We denote the experimental outcome of measuring $\Pi_{i\mu}$ in the $\p{i,j}$-th copy of $\rho$ by $\tilde\Pi_{i\mu}^j\in\set{0,1}$.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Covariance of errors in the spin-projection basis}
+\subsection{Errors in the spin-projection basis}
 
@@
 \begin{align}
-  \bbk{\epsilon_\O} = \Bbk{\tilde\O - \bk{\O}}
+  \bbk{\epsilon_\O} = \bbk{\tilde\O - \bk{\O}}
   = \bk{\O - \bk{\O}}
@@
 \end{align}
-where the double brackets $\bbk{\cdot}$ to denote statistical averaging over experimental trials.
+where the double brackets $\bbk{\cdot}$ to denote statistical averaging over experimental trials that estimate $\bk{\O}$.
 However, the covariance between statistical errors $\epsilon_\O$ and $\epsilon_\Q$ on the empirical estimates $\tilde\O$ and $\tilde\Q$ of observables $\O$ and $\Q$ is
@@
   \bbk{\epsilon_\O \epsilon_\Q}
-  = \Bbk{\p{\tilde\O - \bk{\O}} \p{\tilde\Q - \bk{\Q}}}
+  = \bbk{\p{\tilde\O - \bk{\O}} \p{\tilde\Q - \bk{\Q}}}
   = \bk{\p{\O - \bk{\O}} \p{\Q - \bk{\Q}}}
@@
 \end{align}
-in the empirical estimate of $\bk{\Pi_{i\mu}}_{\rho}$ or $\bk{\bar\Pi_{i\mu}}_{\rho^{\otimes N}}$, and use Eq.~\eqref{eq:mean_proj} to expand
+in the empirical estimate of $\bk{\Pi_{i\mu}}_{\rho}$, and use Eq.~\eqref{eq:mean_proj} to expand
 \begin{align}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Variance of errors in the polarization operator basis}
+\subsection{Errors in the polarization operator basis}
 
@@
   t_{\ell\mu} \equiv \bk{\mu|T_{\ell,0}|\mu}
-  = \sqrt{\f{2\ell+1}{d}} \bk{s\mu;\ell|s\mu},
-\end{align}
-and likewise $\tilde T_{i\ell} = \sum_\mu t_{\ell\mu} \tilde\Pi_{i\mu}$.
+  = \sqrt{\f{2\ell+1}{d}} \bk{s\mu;\ell,0|s\mu},
+\end{align}
+and likewise $\tilde T_{i\ell} \equiv \sum_\mu t_{\ell\mu} \tilde\Pi_{i\mu}$.
 The covariance between errors in the polarization operator basis is then
@@
 
-To this end, we define the probability $p^i_\mu\equiv\bk{\Pi_{i\mu}}_\rho$, collect these probabilities into the classical probability distribution $p^i = \sum_\mu p^i_\mu \ket{\mu}$, and define the vector $t_\ell \equiv \sum_\mu t_{\ell\mu} \ket{\mu}$ of all $t_{\ell\mu}\equiv\bk{\mu|T_{\ell,0}|\mu}$.
+To this end, we define the probability $p^i_\mu\equiv\bk{\Pi_{i\mu}}_\rho$, collect these probabilities into the classical probability distribution $p^i = \sum_\mu p^i_\mu \ket{\mu}$, and define the vector $t_\ell \equiv \sum_\mu t_{\ell\mu} \ket{\mu}$.
 We then observe that

### 2020-12-18 sun_phases/spin_qudit_tomo.tex
tweaks to spin qudit tomography paper, prior to major modifications

@@
 
-\usepackage[inline]{enumitem} % in-line lists and \setlist{} (below)
+\usepackage[inline]{enumitem} % in-line lists
 \setlist[enumerate,1]{label={(\roman*)}} % default in-line numbering
@@
 The central importance of quantum state tomography has led to the development of techniques based on least-squares inversion \cite{opatrny1997leastsquares}, linear regression \cite{qi2013quantum}, maximum likelihood estimation \cite{teo2011quantumstate, smolin2012efficient}, Bayesian inference \cite{huszar2012adaptive, ferrie2014quantum, granade2016practical}, compressed sensing \cite{gross2010quantum, kalev2015quantum}, neural networks \cite{torlai2018neuralnetwork}, among others too numerous to exhaustively list here.
-% todo: fix citation for teo2011
 These techniques are typically developed in a general information-theoretic setting, and make minimal assumptions about the physical medium of a quantum state.
@@
 
-As a special case, the phase-space representation $\rho^{\t{PS}}$ of a qudit state $\rho$ is commonly known as its Husimi distribution.
+As a special case, the phase-space representation $\rho^{\t{PS}}$ of a spin qudit state $\rho$ is commonly known as its Husimi distribution.
 Performing tomography on an unknown qudit state $\rho$ is therefore equivalent to reconstructing the unknown distribution $\rho^{\t{PS}}$ on the sphere.
@@
 
-Observing that $S_\z=T_{1,0}$ and $S_\pm\propto T_{1,\pm1}$, we can use the product operator expansion of the polarization operators (see Appendix \ref{sec:trans_prod}), properties of Clebsch-Gordan coefficients, properties of Wigner $6$-$j$ symbols, and a computer algebra system to simplify the commutators
+Observing that $S_\z=T_{1,0}$ and $S_\pm\propto T_{1,\pm1}$, we can use the operator product expansion of the polarization operators (see Appendix \ref{sec:trans_prod}), the properties of Clebsch-Gordan coefficients, the properties of Wigner $6$-$j$ symbols, and a computer algebra system to simplify the commutators
 \begin{align}
@@
 We leave this minimization problem to future work, and note that solving it will likely require making use of the positivity conditions derived in Ref.~\cite{kryszewski2006positivity}.
-A loose lower bound on $\norm{\rho-\sigma_V^\star}$ can be found by minimization under the constraint $\norm{\rho}\le1$ (or $\bk{\rho_\ell|\rho_\ell}\le1$), which may provide a tighter upper bound on $\E_V\p{\rho}$ than that in Eq.~\eqref{eq:bound} of the main text.
+A loose lower bound on $\norm{\rho-\sigma_V^\star}$ can be found by minimization under the constraint $\norm{\rho}\le1$, which may provide a tighter upper bound on $\E_V\p{\rho}$ than that in Eq.~\eqref{eq:bound} of the main text.
 

### 2021-03-08 sun_phases/figures/sketches/many_int.tex
adding materials for SU(n) spin model paper

\documentclass[preview,varwidth]{standalone}
\input{preamble.tex}

\begin{document}
\begin{tikzpicture}
  % energy levels
  \coordinate (bl) at (0,0);
  \coordinate (tl) at ($ (bl) + (0,3em) $);
  \coordinate (br) at ($ (bl) + (5em,0) $);
  \coordinate (tr) at ($ (tl) + (5em,0) $);
  \coordinate (bc) at ($ (bl)!.5!(br) $);
  \coordinate (tc) at ($ (tl)!.5!(tr) $);
  \draw[thick] (bl) -- (br);
  \draw[thick] (tl) -- (tr);
  \coordinate (bg) at ($ (bl)!.5!(br) + (0,.2em) $);
  \coordinate (tg) at ($ (tl)!.5!(tr) - (0,.2em) $);
  \draw[<->] (bg) -- (tg) node[pos=.5,right] {$U$};
  % spin sketches
  \node[right, xshift=.3em] at (tr) {\includegraphics[scale=0.7]{many_diff.pdf}};
  \node[right, xshift=.3em] at (br) {\includegraphics[scale=0.7]{many_same.pdf}};
  % state text
  \node[above,align=center] at (tc) {asymmetric};
  \node[below,align=center] at (bc) {permutationally \\[-.3em] symmetric};
\end{tikzpicture}
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2021-03-08 sun_phases/sun_phases.tex
adding materials for SU(n) spin model paper

\documentclass[nofootinbib,twocolumn]{revtex4-2}

%%% linking references
\usepackage[dvipsnames]{xcolor}
\usepackage{hyperref}
\hypersetup{
  breaklinks=true,
  colorlinks=true,
  allcolors=BlueViolet,
}

%%% symbols, notations, etc.
\usepackage{physics,braket,bm,amssymb} % physics and math
\renewcommand{\t}{\text} % text in math mode
\newcommand{\f}[2]{\dfrac{#1}{#2}} % shorthand for fractions
\newcommand{\p}[1]{\left(#1\right)} % parenthesis
\renewcommand{\sp}[1]{\left[#1\right]} % square parenthesis
\renewcommand{\set}[1]{\left\{#1\right\}} % curly parenthesis
\newcommand{\bk}{\Braket} % shorthand for braket notation
\renewcommand{\v}{\bm} % bold vectors
\renewcommand{\c}{\cdot} % inner product
\renewcommand{\d}{\partial} % partial derivative
\renewcommand{\dd}{\text{d}} % for infinitesimals
\renewcommand{\i}{\mathrm{i}\mkern1mu} % imaginary unit

% double angle brackets
\newcommand{\bbk}[1]{\langle\!\langle #1 \rangle\!\rangle}
\newcommand{\Bbk}[1]
{\left\langle\!\!\left\langle #1 \right\rangle\!\!\right\rangle}

\usepackage{dsfont} % for identity operator
\newcommand{\1}{\mathds{1}}

% more shorthands
\newcommand{\up}{\uparrow}
\newcommand{\dn}{\downarrow}
\newcommand{\x}{\text{x}}
\newcommand{\y}{\text{y}}
\newcommand{\z}{\text{z}}
\newcommand{\g}{\text{g}}
\newcommand{\e}{\text{e}}
\newcommand{\su}{\mathfrak{su}}
\renewcommand{\SS}{\mathbb{S}}

%%% figures
\usepackage{graphicx} % for figures
\graphicspath{{./figures/}} % set path for all figures

% for text markup
\newcommand{\red}[1]{{\color{red} #1}}

% use alphanumeric tags for footnotes
\renewcommand*{\thefootnote}{\alph{footnote}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\title{SU($n$) spin models in cold atomic systems}
\author{Michael A. Perlin}
% \orcid{https://orcid.org/0000-0002-9316-1596}
\email{mika.perlin@gmail.com}
\author{Diego Barbarena}
\author{Robert J.~Lewis-Swan}
% \orcid{https://orcid.org/0000-0002-0201-281X}
\author{Ana Maria Rey}
% \orcid{http://orcid.org/0000-0001-7176-9413}
\affiliation{JILA, National Institute of Standards and Technology and University of Colorado, 440 UCB, Boulder, Colorado 80309, USA}
\affiliation{Center for Theory of Quantum Matter, University of Colorado, Boulder, CO, 80309, USA}

\date{\today}

\begin{abstract}
  Abstract.
\end{abstract}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:intro}

SU($n$) symmetries play an important role in physics.
Underpinning much of high energy physics, the SU($n$) gauge theory known as Yang-Mills theory is central to our understanding of the electroweak and strong forces.
Extensions of Yang-Mills and SU($n$) symmetry feature in the most well-studied examples of holographic duality \cite{maldacena1999largen} and the connection between entanglement and gravity \cite{ryu2006holographic} through the anti-de Sitter/conformal field theory (AdS/CFT) correspondence.
In a condensed matter setting, SU(2) appears ubiquitously as a symmetry of the Hubbard model, with important consequences for the study of quantum magnetism and high temperature superconductivity \cite{lee2006doping}.
The extension of SU(2) Hubbard and spin models to SU($n$) has led to predictions of exotic phases of matter such as valence bond solids \cite{read1989valencebond, rokhsar1990quadratic, kaul2012lattice, hermele2011topological} and chiral spin liquids \cite{hermele2009mott, hermele2011topological, chen2016syntheticgaugefield, nataf2016chiral}, as well as the potential to perform universal topological quantum computation \cite{freedman2004class, nayak2008nonabelian} and other phenomena \cite{nataf2014exact, nataf2016exact}.
Furthermore, the consideration of disordered SU($n$) spin models has opened analytically tractable avenues for studying quantum chaos and information scrambling \cite{sachdev1993gapless, bentsen2019integrable}.

The tremendous theoretical significance of SU($n$) symmetries makes it all the more exciting that they appear naturally in experimental cold atomic, molecular, and optical (AMO) platforms \cite{gorshkov2010twoorbital, beverland2016realizing, cazalilla2014ultracold, taie2012su, hofrichter2016direct, cappellini2014direct, scazza2014observation, zhang2014spectroscopic, goban2018emergence, perlin2019effective} with exquisite degrees of microscopic control.
In the simplest example, this symmetry arises through the independence of atomic orbital and interaction parameters on the $n$ nuclear spin states of alkaline-earth(-like) atoms, with e.g.~$n=10$ in the case of ${}^{87}$Sr \cite{cazalilla2014ultracold}.
As a result, experiments can directly probe the SU($n$) Hubbard model, leading to experimental observations of SU($n$) Hubbard phases and phase transitions \cite{taie2012su, hofrichter2016direct}, two-orbital SU($n$) magnetism \cite{cappellini2014direct, scazza2014observation, zhang2014spectroscopic}, and multi-body SU($n$)-symmetric interactions \cite{goban2018emergence, perlin2019effective}.
In the spirit of quantum simulation, further investigations of this sort can play an important role in understanding SU($n$) symmetries' consequences for fundamental questions in physics, as well as their practical use in technological applications.
SU(2)-symmetric spin interactions, for example, can enable the development of quantum sensors that use entanglement to surpass classical limits on measurement precision \cite{he2019engineering, perlin2020spin}.
The exciting prospect of similarly exploiting larger SU($n$) symmetries to achieve a technological advantage is still an unexplored avenue of research with enormous potential.

\red{[summarize paper and findings (phases etc.)]}
% (un)interested reader can ...
% main results in/below Eqs. ...

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{From lattice fermions to an SU($n$) spin model}
\label{sec:spin_model}

Here we review the realization of a collective SU($n$) spin model in a system of ultracold alkaline-earth(-like) atoms on an optical lattice.
Without external driving fields, the evolution of such atoms in their electronic ground state is governed by single-body kinetic and two-body interaction Hamiltonians of the form
\begin{align}
  H_{\t{kin}}
  &= -\f{J}{2} \sum_{\bk{j,j'},\mu} c_{j\mu}^\dag c_{j'\mu} + \t{h.c.}, \\
  H_{\t{int}}
  &= \f{U_0}{2} \sum_{j,\mu,\nu}
  c_{j\mu}^\dag c_{j\mu} c_{j\nu}^\dag c_{j\nu},
\end{align}
where $\bk{j,j'}$ denotes neighboring lattice sites $j$ and $j'$; $\mu,\nu\in\set{s,s-1,\cdots,-s}$ index orthogonal spin states of a spin-$s$ nucleus, with $s=\frac{n-1}{2}$ (e.g.~$s=\frac{9}{2}$ in the case of ${}^{87}$Sr with $10$ nuclear spin states); $c_{j\mu}$ is a fermionic annihilation operator, $J$ is a tunneling amplitude; and $U_0$ is a two-body on-site interaction energy.
In the present work, we neglect inter-site interactions and interaction-assisted hopping, which may become relevant in a sufficiently shallow lattice, namely when $J\gtrsim E_{\t{R}}$, where $E_{\t{R}}$ is the atomic lattice recoil energy.
For simplicity, we now assume a periodic lattice of $L$ sites, and expand the on-site fermionic operators in terms of operators addressing (quasi-)momentum modes $q$, $c_{j\mu}=\frac1{\sqrt{L}}\sum_k e^{-\i q\c j} c_{q\mu}$, finding that
\begin{align}
  H_{\t{kin}}
  &= -J\sum_q \cos\p{q} c_{q\mu}^\dag c_{q\mu}, \\
  H_{\t{int}}
  &= \f{U}{N} \sum_{k,\ell,p,q,\mu,\nu}
  c_{k\mu}^\dag c_{\ell\mu} c_{p\nu}^\dag c_{q\nu}
  \times \delta_{k+p,\ell+q},
\end{align}
where $N$ is the total number of atoms on the lattice, we define $U\equiv \frac{U_0}{2}\times\frac{N}{L}$ for convenience, $\delta_{k+p,\ell+q}=1$ if $k+p=\ell+q$ and $0$ otherwise (enforcing conservation of momentum), and we work in units with lattice spacing $a=1$.

If the interaction energy $U_0$ is smaller than the single-particle bandwidth $\sim J$, then the mode-changing collisions in $H_{\t{int}}$ become off-resonant, motivating the frozen-mode approximation $\set{k,p}=\set{\ell,q}$.
In this case, the terms in $H_{\t{int}}$ with $k=\ell$ and $p=q$ become proportional to the total atom number $N$, and can be neglected in the absence of coherent atom number fluctuations.
Defining the spin operators $s_{\mu\nu q}\equiv c_{q\mu}^\dag c_{q\nu}$, the remaining terms of the kinetic and interaction Hamiltonians become
\begin{align}
  H_{\t{kin}} &= -J\sum_q \cos\p{q} s_{\mu\mu q}, \\
  H_{\t{int}} &= -\f{U}{N} \sum_{p,q,\mu,\nu} s_{\mu\nu p} s_{\nu\mu q}.
  \label{eq:H_int_start}
\end{align}
Throughout this work, we will assume that atomic modes are singly-occupied, e.g.~due to the initialization of a spin-polarized state with one atom per lattice site, in which multiple occupation of an atomic mode is forbidden by Pauli exclusion.
In this case, we can neglect the fermionic origins of $H_{\t{int}}$ and instead treat our system as a bona fide spin model.

To further simplify notation and write the interaction Hamiltonian in a form suggestively similar to SU(2) spin alignment, we now construct the operator-valued spin matrix
\begin{align}
  \v s_q \equiv \sum_{\mu,\nu} s_{\mu\nu q} \op{\mu}{\nu},
\end{align}
and for any pair of such operator-valued matrices $\v A,\v B$, we define the inner product
\begin{align}
  \v A \c \v B \equiv \Tr\p{\v A^\dag\v B}
  = \sum_{\mu,\nu} A_{\mu\nu}^\dag B_{\mu\nu},
\end{align}
where the trace is performed only over the auxiliary matrix degrees of freedom.
These definitions allow us to write the spin Hamiltonian in Eq.~\eqref{eq:H_int_start} as
\begin{align}
  H_{\t{int}} = -\f{U}{N} \sum_{p,q} \v s_p\c\v s_q
  = -\f{U}{N}\v S\c\v S,
  &&
  \v S \equiv \sum_q \v s_q.
  \label{eq:H_int}
\end{align}
Here $\v S$ is a collective spin matrix, analogous to the collective spin vector $\vec S=\p{S_\x,S_\y,S_\z}$ in the case of SU(2) \cite{he2019engineering}.

We now discuss the spin Hamiltonian $H_{\t{int}}$ in Eq.~\eqref{eq:H_int}.
The operators $\v s_p\c\v s_q$ in Eq.~\eqref{eq:H_int} simply swap the nuclear spin states of two atoms pinned to modes $p,q$.
The term $\v s_p\c\v s_q$ thereby assigns a definite energy of $+1$ ($-1$) to a pair of spins that are symmetric (anti-symmetric) under exchange.
It is in this sense that $\v s_p\c\v s_q$ is analogous to spin alignment in the case of SU(2), which assigns distinct definite energies to anti-symmetric spin-0 singlet $\ket{\up\dn}-\ket{\dn\up}$ and the symmetric spin-1 triplets $\set{\ket{\up\up},\ket{\dn\dn},\ket{\up\dn}+\ket{\dn\up}}$.

By summing over all pair-wise exchange terms $\v s_p\c\v s_q$, the interaction Hamiltonian $H_{\t{int}}$ energetically enforces a permutational symmetry among all spins, opening an energy gap $U$ between the (degenerate) manifold of all permutationally-symmetric (PS) states and the orthogonal complement of excited (e.g.~spin-wave) states that break permutational symmetry.
In the case of SU(2), the PS manifold is precisely the Dicke manifold of collective states $\ket{m_\z}$ with total spin $S=\frac{N}{2}$ and definite spin projection $m_\z\in\set{S,S-1,\cdots,-S}$ onto a fixed quantization axis.
Equivalently, Dicke states $\ket{m_\z}=\ket{m_\up,m_\dn}$ can be labeled by a definite number of spins $m_\up=S+m_\z$ ($m_\dn=S-m_\z$) pointing up (down) along the spin quantization axis, with $m_\up+m_\dn=N$.
In the general case of SU($n$), the PS manifold is similarly spanned by states $\ket{m_s,m_{s-1},\cdots,m_{-s}}$ with a definite number $m_\mu$ of spins in state $\mu$, and $\sum_\mu m_\mu=N$.
The dimension of the PS manifold, or the ground-state degeneracy of $H_{\t{int}}$, is equal to the number of ways of assigning $N$ identical spins to $n$ distinct internal states, or ${N+n-1 \choose n-1} \sim N^{n-1}$.

\begin{figure}
\centering
\includegraphics[width=\linewidth]{sketches/spin_model.pdf}
\caption{
(a) Ultracold atoms on a lattice can tunnel between neighboring lattice sites at a rate $J$, and locally repel each other with interaction energy $U_0$.
(b,c) When the interaction energy $U_0$ is smaller than the single-particle bandwidth $\sim J$, the frozen-mode approximation enables writing the interaction Hamiltonian as a spin model consisting of exchange terms $\v s_p\c\v s_q$, which swap the states two spins pinned to momentum modes $p,q$.
(d) Interactions thereby open an energy gap $U$ between the manifold of permutationally-symmetric states, and the orthogonal complement of states that break permutational symmetry.
}
\label{fig:spin_model}
\end{figure}

Finally, we take a moment to discuss individual $n$-level spins.
Whenever possible, it is desirable to find geometric representations of abstract mathematical objects, such as the state of a quantum system.
Geometric representations facilitate the development of understanding and intuitions that can be difficult to extract from algebraic expressions alone.
For this reason, the state of a two-level spin, or a qubit, is commonly represented by a point on (or within) the Bloch sphere.
More generally, the state $\ket\psi$ of a $n$-level spin can be represented by a probability distribution\footnote{Strictly speaking, in order to make $Q_\psi$ and $Q_\rho$ probability distributions one has to divide them by a normalization factor $\frac{4\pi}{d}$.
We omit this normalization factor for convenience.} $Q_\psi$ on the sphere $\SS^2$, with the value $Q_\psi\p{\v v}$ at a point $\v v\in\SS^2$ equal to the overlap of $\ket\psi$ with a pure state $\ket{\v v}$ polarized along $\v v$: $Q_\psi\equiv\abs{\bk{\v v|\psi}}^2$ (see Figure \ref{fig:spin_dist}).
In the case of a mixed state $\rho$, this distribution is defined by $Q_\rho\equiv\bk{\v v|\rho|\v v}$.
The dimension $n$ is reflected in $Q_\rho$ by the fact that it can be decomposed into spherical harmonics $Y_{\ell m}$ with degree $\ell<n$ (that is, $Q_\rho$ is ``low resolution'' or ``band-limited'' at degree $n$) \cite{perlin2020qudit}, as well as in the normalization $\int_{\v v\in\SS^2}\dd^2\v v\,Q_\rho\p{\v v}=\frac{4\pi}{d}$ (independent of $\rho$).

\begin{figure}
\centering
\includegraphics[width=\linewidth]{sketches/qudit_rep.pdf}
\caption{
Whereas the state of a two-level spin (qubit) can be represented by a point on (or inside) the Bloch sphere, the state of a $n$-level spin is more generally represented by a probability distribution on the sphere, decomposable into spherical harmonics $Y_{\ell m}$ of degree $\ell<n$.
The distribution shown for $n=10$ corresponds to a Haar-random pure state.
}
\label{fig:spin_dist}
\end{figure}

In fact, $Q_\rho$ is essentially the Husimi-$Q$ function commonly used (e.g.~in the spin squeezing community \cite{ma2011quantum}) to represent collective states of spin-$\frac{1}{2}$ particles in the Dicke manifold.
This correspondence is made precise by identifying the Hilbert space of a single $n$-level spin with the Dicke manifold of $n-1$ spin-$\frac{1}{2}$ particles.
The Hilbert space of a 10-level (spin-$\frac92$) nucleus of a ${}^{87}$Sr atom, for example, can be thought of as the Dicke manifold of 9 ``excess'' spins of its underlying protons and neutrons.
This representation of a $n$-level spin is a useful picture to keep in mind, and will be used freely throughout this work.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Spin-orbit coupling}
\label{sec:SOC}

We wish to study the dynamical behaviors of experimentally realizable SU($n$) spin models.
In order to ensure validity of the spin Hamiltonian $H_{\t{int}}$ in Eq.~\eqref{eq:H_int}, we consider initial preparation of a spin-polarized state, which has the added benefit of being simple to prepare experimentally.
However, a spin polarized state is an eigenstate of $H_{\t{int}}$, so we need additional ingredients to induce nontrivial behavior.
Here, we consider the effect of spin-orbit coupling (SOC) induced by external driving fields.

Before discussing SOC in the context $n$-level fermions on a lattice, we first consider the well-studied case of SOC for two-level atoms in one spatial dimension \cite{wall2016synthetic, bromley2018dynamics}.
In this case, SOC is induced by an external driving field that imprints a phase $e^{-\i\phi j}$ on lattice site $j$, or equivalently imparts a momentum kick $q\to q+\phi$, upon the absorption of a photon:
\begin{align}
  H_{\t{drive}}^{(\phi)}
  = \f{\Omega}{2} \sum_q c_{q+\phi,\up}^\dag c_{q,\dn} + \t{h.c.}.
  \label{eq:drive_2}
\end{align}
Identifying a numerical spin index $\mu=+\frac12$ ($-\frac12$) with the state $\up$ ($\dn$), this drive Hamiltonian can be diagonalized in its momentum index $q$ by the gauge transformation $c_{q\mu}^\dag\to c_{q-\mu\phi,\mu}^\dag$ (equivalently $c_{j\mu}^\dag\to e^{\i\phi\mu j} c_{j\mu}^\dag$), which takes
\begin{align}
  H_{\t{drive}}^{(\phi)} \to H_{\t{drive}} \equiv \Omega S_\x,
  &&
  S_\x \equiv \sum_q s_{\x,q},
  \label{eq:drive_trans}
\end{align}
where $S_\x$ is a collective spin-$x$ operator, and $s_{\x,q}=\frac12 c_{q,\up}^\dag c_{q,\dn} + \t{h.c.}$ for two-level spins.

The drive in Eq.~\eqref{eq:drive_2} is typically implemented by coupling the electronic states of an atom with an external laser \cite{wall2016synthetic, bromley2018dynamics}.
Whereas the coupling between laser fields and nuclear spins is too weak to drive nuclear spin transitions directly, it will nonetheless be useful to consider the the $n$-level (spin-$s$) generalization of Eq.~\eqref{eq:drive_2}.
This generalization will capture the important features of SOC for spin-$s$ nuclei, while avoiding complications that we defer to Section \ref{sec:controls}.

The spin-$s$ generalization of Eq.~\eqref{eq:drive_2} is
\begin{align}
  H_{\t{drive}}^{(\phi)} = \Omega_0 \sum_{q,\mu} \bk{s,\mu;1,1|s,\mu+1}
  c_{q+\phi,\mu+1}^\dag c_{q\mu} + \t{h.c.},
\end{align}
where $\mu\in\set{s,s-1,\cdots,-s}$, and $\bk{s,\mu;1,1|s,\mu+1}$ is a Clebsch-Gordan coefficient.
An identical gauge gauge transformation $c_{q\mu}^\dag\to c_{q-\mu\phi,\mu}^\dag$ again takes $H_{\t{drive}}^{(\phi)}\to \Omega S_\x$ as in Eq.~\eqref{eq:drive_trans}, where now
\begin{align}
  \Omega = \f{\Omega_0}{\xi},
  &&
  \xi \equiv -\sqrt{\f{s(s+1)}{2}},
\end{align}
and
\begin{align}
  s_{\x,q} \equiv \xi \sum_\mu \bk{s,\mu;1,1|s,\mu+1} c_{q,\mu+1}^\dag c_{q\mu} + \t{h.c.}.
  \label{eq:spin_x_op}
\end{align}
is a $n$-level spin-$x$ operator for spin $q$, such that $e^{-\i\theta s_\x}$ rotates the state of spin $q$ about the $x$ axis by an angle $\theta$.
Despite complicating details, the change from two-level to $n$-level spins thus results in a drive Hamiltonian with an identical physical interpretation: after moving into an appropriate ``gauged frame'', this drive simply rotates all spins about the $x$ axis at a rate $\Omega$.

\begin{figure}
\centering
\includegraphics{sketches/soc_panels.pdf}
\caption{
Spin-orbit coupling for dimension-2 (spin-$\frac12$; a,c) and dimension-4 (spin-$\frac32$; b,d) spins.
Colors indicate different nuclear spin projections $\mu$.
In the ``lab frame'' (a,b), kinetic energy is insensitive to spin, but the drive imparts a momentum kick $q\to q+\phi$ into any atom that absorbs a photon.
Changing to the ``gauged frame'' (c,d) makes the drive diagonal in momentum, but comes at the cost of making kinetic energy spin-dependent.
}
\label{fig:soc_panels}
\end{figure}

Spin-orbit coupling cannot be ``gauged away'' entirely, of course.
Making a gauge transformation to simplify the drive comes at the cost of making kinetic energy spin dependent, taking
\begin{align}
  H_{\t{kin}} \to H_{\t{kin}}^{(\phi)}
  \equiv -J \sum_q \cos\p{q+\mu\phi} s_{\mu\mu q},
\end{align}
as visualized in Figure \ref{fig:soc_panels}.
To better interpret this Hamiltonian, we can write it in the simplified form
\begin{align}
  H_{\t{kin}}^{(\phi)}
  = -J \sum_q
  \sp{\cos\p{q} w_{+,q}^{(\phi)} - \sin\p{q} w_{+,q}^{(\phi)}},
\end{align}
where
\begin{align}
  w_{+,q}^{(\phi)} \equiv \cos\p{\mu\phi} s_{\mu\mu q},
  &&
  w_{-,q}^{(\phi)} \equiv \sin\p{\mu\phi} s_{\mu\mu q}
\end{align}
are respectively even ($+$) and odd ($-$) under spin inversion, $\mu\to-\mu$.
For two-level spins, $w_{+,q}^{(\phi)}$ is proportional to the identity operator, and $w_{+,q}^{(\phi)}=2\sin\p{\phi/2} s_{q,\z}$, where $s_{\z,q}=\sum_\mu \mu\, s_{\mu\mu q}$ is a spin-$z$ operator for spin $q$, so the kinetic Hamiltonian in the gauged frame describes a synthetic inhomogeneous magnetic field:
\begin{align}
  \left. H_{\t{kin}}^{(\phi)} \right|_{d=2}
  = 2\sin\p{\phi/2} J \sum_q \sin\p{q} s_{\z,q}.
\end{align}
When $n>2$, an inhomogeneous magnetic field is likewise recovered in the limit of weak SOC, $s\phi\ll1$, in which case
\begin{align}
  \left. H_{\t{kin}}^{(\phi)} \right|_{s\phi\ll1}
  = J\phi \sum_q \sin\p{q} s_{\z,q} + O\p{(s\phi)^2}.
\end{align}
As the strength of SOC is increased, this Hamiltonian acquires terms with higher powers of $s_{\z,q}$, up to $s_{\z,q}^{n-1}$.

\vspace{3cm}
\red{todo: discuss changes to the interaction Hamiltonian}
\begin{align}
  H_{\t{int}} \to H_{\t{int}}^{(\phi)}
  \equiv -\f{U}{N} \sum_{p,q,\mu,\nu}
  c_{p-\mu\phi,\mu}^\dag c_{p-\nu\phi,\nu}
  c_{q-\nu\phi,\nu}^\dag c_{q-\mu\phi,\mu}
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{External controls, observables, and initial states}
\label{sec:controls}

We now take a closer look at external control fields, which determine the initial states we can prepare and the observables we have access to.
To this end, we consider addressing the atoms with several lasers detuned by $\Delta$ below an electronic ${^1}\t{S}_0\to {^3}\t{P}_0$ transition of individual (alkaline-earth-like) atoms on a one-dimensional lattice with tight transverse confinement.
Each laser is indexed by an axis of propagation, $\v v$, and a drive amplitude $\Omega_{\v v,+}$ ($\Omega_{\v v,-}$) for right-circularly (left-circularly) polarized light.
An electronic ${^1}\t{S}_0\to {^3}\t{P}_0$ transition is thus accompanied by a nuclear spin transition $\mu_{\v v}\to\mu_{\v v}\pm1$, where $\mu_{\v v}$ is the projection of nuclear spin onto the axis $\v v$.
Furthermore, laser $\v v$ will generally imprint a phase $e^{-\i\kappa\v v\c\v\ell j}$ on lattice site $j$, where $\kappa$ is the laser's wavenumber (in units with lattice spacing $a=1$) and $\v\ell$ is a fixed unit vector parallel to the lattice.
Altogether, the multi-laser drive Hamiltonian can be written in the form
\begin{multline}
  H_{\t{drive}}^{\t{full}}
  = \sum_{j,\v v,\sigma} \Omega_{\v v\sigma}
  \p{e^{-\i\kappa\v v\c\v\ell j} s_{\v v\sigma j}\otimes\op{\e}{\g}_j + \t{h.c.}} \\
  + \Delta \sum_j \1_j \otimes \op{\e}_j,
  \label{eq:drive_full}
\end{multline}
where $\sigma\in\set{+1,-1}$ indexes a polarization of light; $s_{\v v\sigma j}$ is a $n$-level spin-raising ($\sigma=+1$) or spin-lowering ($\sigma=-1$) operator along axis $\v v$ for atom $j$ (clarified below); $\ket\g_j$ and $\ket\e_j$ respectively denote the ground (${^1}\t{S}_0$) and excited (${^3}\t{P}_0$) electronic states of an atom on site $j$; $\1_j\equiv\sum_\mu s_{\mu\mu j}$ is an identity operator for the spin of atom $j$; and we work in a rotating frame in which the energy difference between ground and excited electronic states is $\Delta$.

To define the spin-raising/lowering operators $s_{\v v\sigma j}$, we first define $s_{\x,j}$ similarly to $s_{\x,q}$ in Eq.~\eqref{eq:spin_x_op} with $q\leftrightarrow j$.
The spin-raising operator $s_{+,j}$ is then twice the spin-raising part of $s_{\x,j}$, just as in the case of SU(2) where $s_\x=\frac12\op{\up}{\dn}+\t{h.c.}$, and $s_{-,j}\equiv s_{+,j}^\dag$.
The operator $s_{\v v,+\,j}$ that raises spin along $\v v$ is acquired by a rotation of $s_{+,j}$ that maps the $z$ axis to $\v v$:
\begin{align}
  s_{\v v,+} \equiv U_{\v v} s_+ U_{\v v}^\dag,
  &&
  U_{\v v} \equiv e^{-\i\alpha_{\v v}s_\z} e^{-\i\beta_{\v v}s_\y},
  \label{eq:spin_rot}
\end{align}
where we have dropped the site index $j$ for brevity, and $s_\y\equiv-\frac\i2s_++\t{h.c.}$.
Note that the operators $(s_\x,s_\y,s_\z)$ form an $\su(2)$ algebra, so given a particular axis $\v v$ one can expand $s_{\v v,+}=\sum_m c_{\v vm} s_m$ with coefficients $c_{\v vm}$ that are independent of the dimension $n$.
In particular, one is always free to set $n=2$ and expand Eq.~\eqref{eq:spin_rot} with spin-$\frac12$ operators (or Pauli matrices) to find the coefficients $c_{\v vm}$.

\begin{figure}
\centering
\includegraphics[width=\linewidth]{sketches/3LD.pdf}
\caption{
Sketch of the three-laser drive used to address nuclear spins on a one-dimensional lattice.
Two counter-propagating lasers with right-circular polarization and amplitudes $\Omega_\pm$ point at an angle $\theta$ to the lattice axis.
A third, linearly polarized laser with amplitude $\Omega_0$ points in a direction orthogonal to both the lattice and the other driving lasers.
Absorbing a photon from the laser with amplitude $\Omega_m$ induces a transition $(\g,\mu)\to(\e,\mu+m)$ for the (electronic, nuclear spin) state of an atom, where nuclear spin is quantized along the axis of propagation for the $m=+1$ laser.
}
\label{fig:3LD}
\end{figure}

In order to make the drive Hamiltonian in Eq.~\eqref{eq:drive_full} more tractable, we orient the lattice along the $z$ axis, $\v\ell=(0,0,1)$, and consider a three-laser drive with a geometry sketched in Figure \ref{fig:3LD}.
In this setup, two counter-propagating right-circularly polarized lasers point in directions at an angle $\theta$ with the lattice axis, $\v v_\pm=\pm(0,-\sin\theta,\cos\theta)$, and a third linearly-polarized laser points along $\v v_0=(1,0,0)$.
Choosing $\v v_+$ as the spin quantization axis, we can define new drive amplitudes $\Omega_0\equiv-(\Omega_{\v v_0,+}+\Omega_{\v v_0,-})$ and $\Omega_\pm\equiv\pm\Omega_{\v v_\pm,+}$, as well as an SOC angle $\phi\equiv\kappa\v v_+\c\v\ell=\kappa\cos\theta$, in terms of which the drive Hamiltonian becomes
\begin{multline}
  H_{\t{3LD}}^{\t{full}}
  = \sum_{j,m} \Omega_m
  \p{e^{-\i m\phi j} s_{mj} \otimes\op{\e}{\g}_j + \t{h.c.}} \\
  + \Delta \sum_j \1_j \otimes \op{\e}_j,
\end{multline}
where $m\in\set{+1,0,-1}$ indexes the laser pointing along $\v v_m$; and $s_{0,j}\equiv s_{\z,j}$ for shorthand.
In the far-detuned limit $\Delta\gg\abs{\Omega_m}$, a second-order perturbative treatment of ${^3}\t{P}_0$ excitations ($\ket\e$) yields an effective drive Hamiltonian that only addresses ground-state nuclear spins.
After a gauge transformation $c_{j\mu}^\dag\to e^{\i\phi\mu j} c_{j\mu}^\dag$ (equivalently $s_{mj}\to e^{\i m\phi j}s_{mj}$), we can then make the simplifying assumption (or requirement) that all $\Omega_m$ are real to write
\begin{align}
  H_{\t{3LD}} = \sum_j H_{\t{3LD},j}^{\t{single}},
  \label{eq:drive_all}
\end{align}
where $H_{\t{3LD},j}^{\t{single}}$ denotes the action of $H_{\t{3LD}}^{\t{single}}$ on spin $j$:
\begin{multline}
  H_{\t{3LD}}^{\t{single}}
  = \tilde\Omega_+ \tilde\Omega_- s_\z
  + \tilde\Omega_0 \tilde\Omega_- s_\x
  + \tilde\Omega_0 \tilde\Omega_+ \p{s_\z s_\x  + s_\x s_\z} \\
  - \tilde\Omega_0^2 s_\z^2 - \tilde\Omega_+^2 s_\x^2
  - \tilde\Omega_-^2 s_\y^2,
  \label{eq:drive_single}
\end{multline}
with
\begin{align}
  \tilde\Omega_0 \equiv -\f{\Omega_0}{\sqrt\Delta},
  &&
  \tilde\Omega_\pm \equiv \f{\Omega_+\pm\Omega_-}{\sqrt\Delta}.
\end{align}

\begin{table}
\centering
\caption{
Drive Hamiltonians (left column) that can be implemented with different amplitude-matching conditions (right column), some of which are specified by a sign $\sigma\in\set{+1,-1}$.
The drives shown here are equal to that Eq.~\eqref{eq:drive_single} up to a possible energy shift of $s_\x^2+s_\y^2+s_\z^2=s(s+1)$, and come in mutually commuting pairs: a drive with $\Omega_m=1$ and $\Omega_n=0$ for both $n\ne m$ commutes with the drive in which both $\abs*{\Omega_n}=1$ but $\Omega_m=0$.
}
\vspace{.5em}
\begin{tabular}{c|c}
  $H_{\t{drive}}^{\t{single}}$
  & $(\tilde\Omega_0,\tilde\Omega_+,\tilde\Omega_-)$
  \\ \hline\hline
  $-s_\z^2$ & $\p{1,0,0}$
  \\ \hline
   $-s_\x^2$ & $\p{0,1,0}$
  \\ \hline
  $-s_\y^2$ & $\p{0,0,1}$
  \\ \hline
  $\sigma s_\z + s_\z^2$ & $\p{0,1,\sigma}$
  \\ \hline
  $\sigma s_\x + s_\x^2$ & $\p{1,0,\sigma}$
  \\ \hline
  $\sigma\p{s_\z s_\x+s_\x s_\z} + s_\y^2$ & $\p{1,\sigma,0}$
  \\ \hline
  $\pm s_\z \pm \sigma s_\x + \sigma \p{s_\z s_\x + s_\x s_\z}$
  & $\p{1,\sigma,\pm\sigma}$
\end{tabular}
\label{tab:drives}
\end{table}

Eqs.~\eqref{eq:drive_all} and \eqref{eq:drive_single} are key results of this section.
There are three important observations to make about these results.
First, the fact that $H_{\t{3LD}}$ acts identically on all spins means we can freely replace the site index $j$ with a momentum index $q$ (as can be verified by substituting $c_{j\mu}=\frac1{\sqrt{L}}\sum_k e^{-\i q\c j} c_{q\mu}$), which is important to ensure that this drive addresses the same spin degrees of freedom as the final Hamiltonians in Section \ref{sec:SOC}.
Second, each of $\tilde\Omega_0,\tilde\Omega_+,\tilde\Omega_-$ can be tuned independently by changing the amplitudes of the driving lasers; some particular Hamiltonians for specific values of these amplitudes are shown in Table \ref{tab:drives}.
Third, due to the appearance of mutually commuting pairs of Hamiltonians in Table \ref{tab:drives}, specifically $-s_\alpha^2$ and $\pm s_\alpha+s_\alpha^2$ for $\alpha\in\set{\z,\x}$, the three-laser drive admits pulse sequences that can implement arbitrary SU(2) (spatial) rotations of the form $e^{-\i\chi\vec n\c\vec s}$, where $\chi$ is a rotation angle, $\vec n$ is a rotation axis, and $\vec s\equiv(s_\x,s_\y,s_\z)$.
The capability to perform arbitrary spatial rotations, together with the capability to measure the number of atoms with spin projection $\mu$ onto a fixed quantization axis, $\bk{S_{\mu\mu}}$ (where $S_{\mu\nu}=\sum_js_{\mu\nu j}$), implies the capability to reconstruct all components of the mean collective spin matrix $\bk{\v S}=\sum_{\mu\nu}\bk{S_{\mu\nu}}\op{\mu}{\nu}$ via spin qudit tomography \cite{newton1968measurability, perlin2020qudit}.

Finally, we comment on initial states that these external controls allow us to prepare.

\red{comment about initial states (``intra-spin correlations'')}

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{A simple, realistic model}
\label{sec:full_model}

% use/simulate sinusoidal dispersion
% initial state in |X>
% repeat analysis with different initial state
% -- look at class of initial states (e.g. back-to-back, tipping angle)

\bibliography{\jobname.bib}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2021-03-08 sun_phases/figures/sketches/make_plots.py

@@ -1,0 +1,517 @@
+#!/usr/bin/env python3
+
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.patches as patches
+from mpl_toolkits.mplot3d import proj3d
+
+preamble = r"""
+\usepackage{braket}
+"""
+
+fontsize = 10
+params = { "font.family" : "serif",
+           "font.serif" : "Computer Modern",
+           "font.size" : fontsize,
+           "axes.titlesize" : fontsize,
+           "axes.labelsize" : fontsize,
+           "text.usetex" : True,
+           "text.latex.preamble" : preamble }
+plt.rcParams.update(params)
+
+### arrow styles and line colors
+
+arrow_style = "head_width=1.5,head_length=4"
+single_kw = dict(arrowstyle = "-|>,"+arrow_style, color="k")
+double_kw = dict(arrowstyle = "<|-|>,"+arrow_style, color="k")
+
+def marker_kw(color = None):
+    return { "marker" : "o",
+             "color" : color,
+             "markeredgecolor" : "k",
+             "markeredgewidth" : 1 }
+
+color_cycle = [ _["color"] for _ in plt.rcParams["axes.prop_cycle"] ]
+
+dn_color = color_cycle[0]
+up_color = color_cycle[1]
+
+def add_arrow(start, end, axis = None, *args, **kw_args):
+    if axis is None: axis = plt.gca()
+    arrow = patches.FancyArrowPatch(start, end, *args, **kw_args)
+    axis.add_patch(arrow)
+
+##########################################################################################
+# lattice diagram
+##########################################################################################
+
+figsize = (2,0.75)
+plt.figure(figsize = figsize)
+x_lim, y_lim = 1.6, 0.5
+x_vals = np.linspace(-x_lim, x_lim, 300)
+plt.plot(x_vals, -y_lim*np.cos(2*np.pi*x_vals), "k")
+
+btm_y, top_y = -y_lim*0.4, y_lim*0.7
+plt.plot([-1], [btm_y], **marker_kw(dn_color))
+plt.plot([+0], [btm_y], **marker_kw(dn_color))
+add_arrow((-1,btm_y+0.08), (+0,btm_y+0.08),
+      connectionstyle="arc3,rad=-0.4", **single_kw)
+plt.annotate(r"$J$", (-0.5, btm_y), ha = "center", va = "center")
[…]

### 2021-03-22 sun_phases/sun_phases.tex
update multilevel spin model paper

@@
 
-\title{SU($n$) spin models in cold atomic systems}
+\title{Multilevel spin models in cold atomic systems}
 \author{Michael A. Perlin}
@@
 
-The tremendous theoretical significance of SU($n$) symmetries makes it all the more exciting that they appear naturally in experimental cold atomic, molecular, and optical (AMO) platforms \cite{gorshkov2010twoorbital, beverland2016realizing, cazalilla2014ultracold, taie2012su, hofrichter2016direct, cappellini2014direct, scazza2014observation, zhang2014spectroscopic, goban2018emergence, perlin2019effective} with exquisite degrees of microscopic control.
+The tremendous theoretical significance of SU($n$) symmetries makes it all the more exciting that they appear naturally in experimental atomic, molecular, and optical (AMO) platforms \cite{gorshkov2010twoorbital, beverland2016realizing, cazalilla2014ultracold, taie2012su, hofrichter2016direct, cappellini2014direct, scazza2014observation, zhang2014spectroscopic, goban2018emergence, perlin2019effective} with exquisite degrees of microscopic control.
 In the simplest example, this symmetry arises through the independence of atomic orbital and interaction parameters on the $n$ nuclear spin states of alkaline-earth(-like) atoms, with e.g.~$n=10$ in the case of ${}^{87}$Sr \cite{cazalilla2014ultracold}.
@@
 In the spirit of quantum simulation, further investigations of this sort can play an important role in understanding SU($n$) symmetries' consequences for fundamental questions in physics, as well as their practical use in technological applications.
-SU(2)-symmetric spin interactions, for example, can enable the development of quantum sensors that use entanglement to surpass classical limits on measurement precision \cite{he2019engineering, perlin2020spin}.
+For example, SU(2)-symmetric spin interactions can be harnessed to develop quantum sensors that surpass classical limits on measurement precision \cite{he2019engineering, perlin2020spin}.
 The exciting prospect of similarly exploiting larger SU($n$) symmetries to achieve a technological advantage is still an unexplored avenue of research with enormous potential.
 
-\red{[summarize paper and findings (phases etc.)]}
+\red{[finish intro, summarize paper]}
 % (un)interested reader can ...
@@
   H_{\t{int}}
-  &= \f{U}{N} \sum_{k,\ell,p,q,\mu,\nu}
+  &= \f{U}{2N} \sum_{k,\ell,p,q,\mu,\nu}
   c_{k\mu}^\dag c_{\ell\mu} c_{p\nu}^\dag c_{q\nu}
   \times \delta_{k+p,\ell+q},
+  \label{eq:H_int_momenta}
 \end{align}
-where $N$ is the total number of atoms on the lattice, we define $U\equiv \frac{U_0}{2}\times\frac{N}{L}$ for convenience, $\delta_{k+p,\ell+q}=1$ if $k+p=\ell+q$ and $0$ otherwise (enforcing conservation of momentum), and we work in units with lattice spacing $a=1$.
+where $N$ is the total number of atoms on the lattice, we define $U\equiv U_0\times N/L$ for convenience, $\delta_{k+p,\ell+q}=1$ if $k+p=\ell+q$ and $0$ otherwise (enforcing conservation of momentum), and we work in units with lattice spacing $a=1$.
 
-If the interaction energy $U_0$ is smaller than the single-particle bandwidth $\sim J$, then the mode-changing collisions in $H_{\t{int}}$ become off-resonant, motivating the frozen-mode approximation $\set{k,p}=\set{\ell,q}$.
-In this case, the terms in $H_{\t{int}}$ with $k=\ell$ and $p=q$ become proportional to the total atom number $N$, and can be neglected in the absence of coherent atom number fluctuations.
+If the interaction energy $U_0$ is smaller than the single-particle bandwidth $\sim J$, then the mode-changing collisions in $H_{\t{int}}$ become off-resonant, motivating the frozen-mode approximation $\set{k,p}=\set{\ell,q}$ (which we assume for the remainder of this work).
+The operator content of terms with $k=\ell$ and $p=q$ takes the form $c_{\ell\mu}^\dag c_{\ell\mu} c_{q\nu}^\dag c_{q\nu}$, so these terms contribute an overall energy $\frac12NU$ that can be neglected in the absence of coherent atom number fluctuations.
 Defining the spin operators $s_{\mu\nu q}\equiv c_{q\mu}^\dag c_{q\nu}$, the remaining terms of the kinetic and interaction Hamiltonians become
 \begin{align}
-  H_{\t{kin}} &= -J\sum_q \cos\p{q} s_{\mu\mu q}, \\
-  H_{\t{int}} &= -\f{U}{N} \sum_{p,q,\mu,\nu} s_{\mu\nu p} s_{\nu\mu q}.
+  H_{\t{kin}} &= -J\sum_q \cos\p{q} s_{\mu\mu q},
+  \label{H_kin_start} \\
+  H_{\t{int}} &= -\f{U}{2N} \sum_{p,q,\mu,\nu} s_{\mu\nu p} s_{\nu\mu q}.
   \label{eq:H_int_start}
@@
 \begin{align}
-  H_{\t{int}} = -\f{U}{N} \sum_{p,q} \v s_p\c\v s_q
-  = -\f{U}{N}\v S\c\v S,
+  H_{\t{int}} = -\f{U}{2N} \sum_{p,q} \v s_p\c\v s_q
+  = -\f{U}{2N}\v S\c\v S,
   &&
@@
 We now discuss the spin Hamiltonian $H_{\t{int}}$ in Eq.~\eqref{eq:H_int}.
-The operators $\v s_p\c\v s_q$ in Eq.~\eqref{eq:H_int} simply swap the nuclear spin states of two atoms pinned to modes $p,q$.
+The operator $\v s_p\c\v s_q$ simply swaps the nuclear spin states of two atoms pinned to modes $p,q$.
 The term $\v s_p\c\v s_q$ thereby assigns a definite energy of $+1$ ($-1$) to a pair of spins that are symmetric (anti-symmetric) under exchange.
-It is in this sense that $\v s_p\c\v s_q$ is analogous to spin alignment in the case of SU(2), which assigns distinct definite energies to anti-symmetric spin-0 singlet $\ket{\up\dn}-\ket{\dn\up}$ and the symmetric spin-1 triplets $\set{\ket{\up\up},\ket{\dn\dn},\ket{\up\dn}+\ket{\dn\up}}$.
+In this sense, $\v s_p\c\v s_q$ is analogous to the enforcement of SU(2) spin alignment, which similarly assigns (distinct) definite energies to anti-symmetric spin-0 singlet $\ket{\up\dn}-\ket{\dn\up}$ and the symmetric spin-1 triplets $\set{\ket{\up\up},\ket{\dn\dn},\ket{\up\dn}+\ket{\dn\up}}$.
 
-By summing over all pair-wise exchange terms $\v s_p\c\v s_q$, the interaction Hamiltonian $H_{\t{int}}$ energetically enforces a permutational symmetry among all spins, opening an energy gap $U$ between the (degenerate) manifold of all permutationally-symmetric (PS) states and the orthogonal complement of excited (e.g.~spin-wave) states that break permutational symmetry.
+By summing over all pair-wise exchange terms $\v s_p\c\v s_q$, the interaction Hamiltonian $H_{\t{int}}$ energetically enforces a permutational symmetry among all spins, opening an energy gap $U$ between the manifold of all permutationally-symmetric (PS) states and the orthogonal complement of excited (e.g.~spin-wave) states that break permutational symmetry.
 In the case of SU(2), the PS manifold is precisely the Dicke manifold of collective states $\ket{m_\z}$ with total spin $S=\frac{N}{2}$ and definite spin projection $m_\z\in\set{S,S-1,\cdots,-S}$ onto a fixed quantization axis.
@@
 Whenever possible, it is desirable to find geometric representations of abstract mathematical objects, such as the state of a quantum system.
-Geometric representations facilitate the development of understanding and intuitions that can be difficult to extract from algebraic expressions alone.
+Geometric representations facilitate the development of understanding and intuitions that can be difficult to extract from algebraic expressions.
 For this reason, the state of a two-level spin, or a qubit, is commonly represented by a point on (or within) the Bloch sphere.
-More generally, the state $\ket\psi$ of a $n$-level spin can be represented by a probability distribution\footnote{Strictly speaking, in order to make $Q_\psi$ and $Q_\rho$ probability distributions one has to divide them by a normalization factor $\frac{4\pi}{d}$.
-We omit this normalization factor for convenience.} $Q_\psi$ on the sphere $\SS^2$, with the value $Q_\psi\p{\v v}$ at a point $\v v\in\SS^2$ equal to the overlap of $\ket\psi$ with a pure state $\ket{\v v}$ polarized along $\v v$: $Q_\psi\equiv\abs{\bk{\v v|\psi}}^2$ (see Figure \ref{fig:spin_dist}).
+More generally, the state $\ket\psi$ of a $n$-level spin can be represented by a probability distribution\footnote{Strictly speaking, in order to make $Q_\psi$ (and $Q_\rho$) a probability distribution, one has to divide it by a normalization factor $\frac{4\pi}{d}$.
+We omit this normalization factor for convenience.} $Q_\psi$ on the sphere $\SS^2$.
+The value $Q_\psi\p{\v v}$ at a point $\v v\in\SS^2$ equal to the overlap of $\ket\psi$ with a pure state $\ket{\v v}$ maximally polarized along $\v v$: $Q_\psi\equiv\abs{\bk{\v v|\psi}}^2$ (see Figure \ref{fig:spin_dist}).
 In the case of a mixed state $\rho$, this distribution is defined by $Q_\rho\equiv\bk{\v v|\rho|\v v}$.
-The dimension $n$ is reflected in $Q_\rho$ by the fact that it can be decomposed into spherical harmonics $Y_{\ell m}$ with degree $\ell<n$ (that is, $Q_\rho$ is ``low resolution'' or ``band-limited'' at degree $n$) \cite{perlin2020qudit}, as well as in the normalization $\int_{\v v\in\SS^2}\dd^2\v v\,Q_\rho\p{\v v}=\frac{4\pi}{d}$ (independent of $\rho$).
+The dimension $n$ is reflected in $Q_\rho$ by the fact that it can be decomposed into spherical harmonics $Y_{\ell m}$ with degree $\ell<n$ (that is, $Q_\rho$ is ``low resolution'' or ``band-limited'' at degree $n$) \cite{perlin2020qudit}, as well as in the normalization $\int_{\v v\in\SS^2}\dd^2\v v\,Q_\rho\p{\v v}=\frac{4\pi}{d}$.
 
@@
 This correspondence is made precise by identifying the Hilbert space of a single $n$-level spin with the Dicke manifold of $n-1$ spin-$\frac{1}{2}$ particles.
-The Hilbert space of a 10-level (spin-$\frac92$) nucleus of a ${}^{87}$Sr atom, for example, can be thought of as the Dicke manifold of 9 ``excess'' spins of its underlying protons and neutrons.
-This representation of a $n$-level spin is a useful picture to keep in mind, and will be used freely throughout this work.
+The Hilbert space of a 10-level (spin-$\frac92$) nucleus of a ${}^{87}$Sr atom, for example, can be thought of as the Dicke manifold of 9 ``excess'' proton and neutron spins.
+This representation of a $n$-level quantum system is a useful picture to keep in mind for angular-momentum-like degrees of freedom, and will be used freely throughout this work.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Spin-orbit coupling}
+\section{Spin-orbit coupling and gauge transformation}
 \label{sec:SOC}
@@
 We wish to study the dynamical behaviors of experimentally realizable SU($n$) spin models.
-In order to ensure validity of the spin Hamiltonian $H_{\t{int}}$ in Eq.~\eqref{eq:H_int}, we consider initial preparation of a spin-polarized state, which has the added benefit of being simple to prepare experimentally.
-However, a spin polarized state is an eigenstate of $H_{\t{int}}$, so we need additional ingredients to induce nontrivial behavior.
-Here, we consider the effect of spin-orbit coupling (SOC) induced by external driving fields.
+In order to ensure validity of the $\v S\c\v S$ spin Hamiltonian in Eq.~\eqref{eq:H_int}, we consider initial preparation of a spin-polarized state\footnote{We remind the reader that validity of the spin model requires that each momentum mode is at most singly-occupied, which is guaranteed in a spin-polarized state due to Pauli blocking.}, which has the added benefit of being simple to prepare experimentally.
+However, a spin polarized state is an eigenstate of both the kinetic and interaction spin Hamiltonians in Section \ref{sec:spin_model}, so we need additional ingredients for nontrivial behavior.
+Here, we consider the addition of spin-orbit coupling (SOC) induced by external driving fields.
 
-Before discussing SOC in the context $n$-level fermions on a lattice, we first consider the well-studied case of SOC for two-level atoms in one spatial dimension \cite{wall2016synthetic, bromley2018dynamics}.
+Before discussing SOC for $n$-level fermions, we first consider the well-studied case of two-level SOC on a one-dimensional lattice \cite{wall2016synthetic, bromley2018dynamics}.
 In this case, SOC is induced by an external driving field that imprints a phase $e^{-\i\phi j}$ on lattice site $j$, or equivalently imparts a momentum kick $q\to q+\phi$, upon the absorption of a photon:
@@
 
-The drive in Eq.~\eqref{eq:drive_2} is typically implemented by coupling the electronic states of an atom with an external laser \cite{wall2016synthetic, bromley2018dynamics}.
-Whereas the coupling between laser fields and nuclear spins is too weak to drive nuclear spin transitions directly, it will nonetheless be useful to consider the the $n$-level (spin-$s$) generalization of Eq.~\eqref{eq:drive_2}.
-This generalization will capture the important features of SOC for spin-$s$ nuclei, while avoiding complications that we defer to Section \ref{sec:controls}.
+The two-level drive in Eq.~\eqref{eq:drive_2} is typically implemented by coupling pseudo-spin electronic states of an atom with an external laser \cite{wall2016synthetic, bromley2018dynamics}.
+Though the coupling between laser fields and nuclear spins is generally too weak to drive nuclear spin transitions directly, it will nonetheless be useful to consider the $n$-level generalization of the two-level drive.
+This generalization will capture the important features of SOC for $n$-level nuclei, while avoiding complications that we defer to Section \ref{sec:controls}.
 
@@
 where $\mu\in\set{s,s-1,\cdots,-s}$, and $\bk{s,\mu;1,1|s,\mu+1}$ is a Clebsch-Gordan coefficient.
-An identical gauge gauge transformation $c_{q\mu}^\dag\to c_{q-\mu\phi,\mu}^\dag$ again takes $H_{\t{drive}}^{(\phi)}\to \Omega S_\x$ as in Eq.~\eqref{eq:drive_trans}, where now
+The gauge gauge transformation $c_{q\mu}^\dag\to c_{q-\mu\phi,\mu}^\dag$ again takes $H_{\t{drive}}^{(\phi)}\to \Omega S_\x$ as in Eq.~\eqref{eq:drive_trans}, where now
 \begin{align}
@@
 \begin{align}
-  s_{\x,q} \equiv \xi \sum_\mu \bk{s,\mu;1,1|s,\mu+1} c_{q,\mu+1}^\dag c_{q\mu} + \t{h.c.}.
+  s_{\x,q} \equiv \xi \sum_\mu \bk{s,\mu;1,1|s,\mu+1} c_{q,\mu+1}^\dag c_{q\mu} + \t{h.c.}
   \label{eq:spin_x_op}
 \end{align}
-is a $n$-level spin-$x$ operator for spin $q$, such that $e^{-\i\theta s_\x}$ rotates the state of spin $q$ about the $x$ axis by an angle $\theta$.
+is an $n$-level spin-$x$ operator for spin $q$, with which $e^{-\i\theta s_{\x,q}}$ rotates the state of spin $q$ about the $x$ axis by an angle $\theta$.
 Despite complicating details, the change from two-level to $n$-level spins thus results in a drive Hamiltonian with an identical physical interpretation: after moving into an appropriate ``gauged frame'', this drive simply rotates all spins about the $x$ axis at a rate $\Omega$.
@@
 \caption{
-Spin-orbit coupling for dimension-2 (spin-$\frac12$; a,c) and dimension-4 (spin-$\frac32$; b,d) spins.
-Colors indicate different nuclear spin projections $\mu$.
-In the ``lab frame'' (a,b), kinetic energy is insensitive to spin, but the drive imparts a momentum kick $q\to q+\phi$ into any atom that absorbs a photon.
-Changing to the ``gauged frame'' (c,d) makes the drive diagonal in momentum, but comes at the cost of making kinetic energy spin-dependent.
+Spin-orbit coupling for 2-level ({\bf a},{\bf c}) and 4-level ({\bf b},{\bf d}) spins.
+Colors indicate different spin projections $\mu$.
+In the ``lab frame'' ({\bf a},{\bf b}), kinetic energy is insensitive to spin, but the drive imparts a momentum kick $q\to q+\phi$ into any atom that absorbs a photon.
+Changing to the ``gauged frame'' ({\bf c},{\bf d}) makes the drive diagonal in momentum, but comes at the cost of making kinetic energy spin-dependent.
 }
@@
 
-Spin-orbit coupling cannot be ``gauged away'' entirely, of course.
+Of course, spin-orbit coupling cannot be ``gauged away'' entirely.
 Making a gauge transformation to simplify the drive comes at the cost of making kinetic energy spin dependent, taking
@@
 \begin{align}
-  w_{+,q}^{(\phi)} \equiv \cos\p{\mu\phi} s_{\mu\mu q},
-  &&
-  w_{-,q}^{(\phi)} \equiv \sin\p{\mu\phi} s_{\mu\mu q}
+  w_{+,q}^{(\phi)} &\equiv \sum_\mu \cos\p{\mu\phi} s_{\mu\mu q}, \\
+  w_{-,q}^{(\phi)} &\equiv \sum_\mu \sin\p{\mu\phi} s_{\mu\mu q}.
 \end{align}
-are respectively even ($+$) and odd ($-$) under spin inversion, $\mu\to-\mu$.
-For two-level spins, $w_{+,q}^{(\phi)}$ is proportional to the identity operator, and $w_{+,q}^{(\phi)}=2\sin\p{\phi/2} s_{q,\z}$, where $s_{\z,q}=\sum_\mu \mu\, s_{\mu\mu q}$ is a spin-$z$ operator for spin $q$, so the kinetic Hamiltonian in the gauged frame describes a synthetic inhomogeneous magnetic field:
+For two-level spins, $w_{+,q}^{(\phi)}$ is proportional to the identity operator, and $w_{+,q}^{(\phi)}=2\sin\p{\phi/2} s_{q,\z}$, where $s_{\z,q}=\sum_\mu \mu\, s_{\mu\mu q}$ is a spin-$z$ operator for spin $q$, so the kinetic Hamiltonian in the gauged frame describes a (synthetic) inhomogeneous magnetic field:
 \begin{align}
@@
 
-\vspace{3cm}
-\red{todo: discuss changes to the interaction Hamiltonian}
+In principle, making the gauge transformation $c_{q\mu}^\dag\to c_{q-\mu\phi,\mu}^\dag$ also transforms the interaction Hamiltonian as
 \begin{align}
   H_{\t{int}} \to H_{\t{int}}^{(\phi)}
-  \equiv -\f{U}{N} \sum_{p,q,\mu,\nu}
+  \equiv -\f{U}{2N} \sum_{p,q,\mu,\nu}
   c_{p-\mu\phi,\mu}^\dag c_{p-\nu\phi,\nu}
-  c_{q-\nu\phi,\nu}^\dag c_{q-\mu\phi,\mu}
+  c_{q-\nu\phi,\nu}^\dag c_{q-\mu\phi,\mu}.
+\end{align}
+However, $H_{\t{int}}^{(\phi)}$ cannot be exactly reduced to a spin Hamiltonian that addresses the same spin degrees of freedom as $H_{\t{kin}}^{(\phi)}$.
+Nonetheless, in the weak SOC limit $s\phi\ll1$ the effect of $H_{\t{int}}^{(\phi)}$ may be approximated by $H_{\t{int}}^{(0)}=H_{\t{int}}$, which has previously been benchmarked for SU(2)-symmetric interactions \cite{he2019engineering, smale2019observation}.
+To ensure that $H_{\t{kin}}^{(\phi)}$ does not become trivial as $\phi\to0$, we can simultaneously take $J\to\infty$ while keeping $J\phi$ constant.
+Altogether, the interacting spin Hamiltonian in the gauged frame becomes
+\begin{align}
+  H_{\t{spin}} = -\f{U}{2N} \v S\c\v S + J\phi \sum_q \sin\p{q} s_{\z,q}.
+  \label{eq:H_spin}
 \end{align}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{External controls, observables, and initial states}
+\section{Control fields, observables, and initial states}
 \label{sec:controls}
 
-We now take a closer look at external control fields, which determine the initial states we can prepare and the observables we have access to.
-To this end, we consider addressing the atoms with several lasers detuned by $\Delta$ below an electronic ${^1}\t{S}_0\to {^3}\t{P}_0$ transition of individual (alkaline-earth-like) atoms on a one-dimensional lattice with tight transverse confinement.
+We now take a closer look at external control fields, which determine the observables we can access and initial states that we can prepare.
+To this end, we consider addressing the atoms with several lasers detuned by $\Delta$ below an electronic ${^1}\t{S}_0\to {^3}\t{P}_0$ transition of individual alkaline-earth-like atoms on a one-dimensional lattice with tight transverse confinement.
 Each laser is indexed by an axis of propagation, $\v v$, and a drive amplitude $\Omega_{\v v,+}$ ($\Omega_{\v v,-}$) for right-circularly (left-circularly) polarized light.
 An electronic ${^1}\t{S}_0\to {^3}\t{P}_0$ transition is thus accompanied by a nuclear spin transition $\mu_{\v v}\to\mu_{\v v}\pm1$, where $\mu_{\v v}$ is the projection of nuclear spin onto the axis $\v v$.
-Furthermore, laser $\v v$ will generally imprint a phase $e^{-\i\kappa\v v\c\v\ell j}$ on lattice site $j$, where $\kappa$ is the laser's wavenumber (in units with lattice spacing $a=1$) and $\v\ell$ is a fixed unit vector parallel to the lattice.
+Furthermore, a laser pointing along $\v v$ will generally imprint a phase $e^{-\i\kappa\v v\c\v\ell j}$ on lattice site $j$, where $\kappa$ is the laser's wavenumber (in units with lattice spacing $a=1$) and $\v\ell$ is a fixed unit vector parallel to the lattice.
 Altogether, the multi-laser drive Hamiltonian can be written in the form
@@
 To define the spin-raising/lowering operators $s_{\v v\sigma j}$, we first define $s_{\x,j}$ similarly to $s_{\x,q}$ in Eq.~\eqref{eq:spin_x_op} with $q\leftrightarrow j$.
-The spin-raising operator $s_{+,j}$ is then twice the spin-raising part of $s_{\x,j}$, just as in the case of SU(2) where $s_\x=\frac12\op{\up}{\dn}+\t{h.c.}$, and $s_{-,j}\equiv s_{+,j}^\dag$.
-The operator $s_{\v v,+\,j}$ that raises spin along $\v v$ is acquired by a rotation of $s_{+,j}$ that maps the $z$ axis to $\v v$:
+The spin-raising operator $s_{+,j}$ is then twice the spin-raising part of $s_{\x,j}$, just as in the case of SU(2) where $s_\x=\frac12\op{\up}{\dn}+\t{h.c.}$, and the spin-lowering operator is $s_{-,j}\equiv s_{+,j}^\dag$.
+The operator $s_{\v v,+\,j}$ that raises spin along $\v v$ is acquired by a rotation of $s_{+,j}$ that maps the north pole to $\v v$:
 \begin{align}
@@
 \end{align}
-where we have dropped the site index $j$ for brevity, and $s_\y\equiv-\frac\i2s_++\t{h.c.}$.
+where we have dropped the site index $j$ for brevity; $\alpha_{\v v}$ and $\beta_{\v v}$ are the azimuthal and polar angles of $\v v$; and $s_\y\equiv-\frac\i2s_++\t{h.c.}$.
 Note that the operators $(s_\x,s_\y,s_\z)$ form an $\su(2)$ algebra, so given a particular axis $\v v$ one can expand $s_{\v v,+}=\sum_m c_{\v vm} s_m$ with coefficients $c_{\v vm}$ that are independent of the dimension $n$.
@@
 In this setup, two counter-propagating right-circularly polarized lasers point in directions at an angle $\theta$ with the lattice axis, $\v v_\pm=\pm(0,-\sin\theta,\cos\theta)$, and a third linearly-polarized laser points along $\v v_0=(1,0,0)$.
-Choosing $\v v_+$ as the spin quantization axis, we can define new drive amplitudes $\Omega_0\equiv-(\Omega_{\v v_0,+}+\Omega_{\v v_0,-})$ and $\Omega_\pm\equiv\pm\Omega_{\v v_\pm,+}$, as well as an SOC angle $\phi\equiv\kappa\v v_+\c\v\ell=\kappa\cos\theta$, in terms of which the drive Hamiltonian becomes
+Choosing $\v v_+$ as the spin quantization axis, we can define new drive amplitudes $\Omega_0\equiv-(\Omega_{\v v_0,+}+\Omega_{\v v_0,-})$ and $\Omega_\pm\equiv\pm\Omega_{\v v_\pm,+}$, as well as a SOC angle $\phi\equiv\kappa\v v_+\c\v\ell=\kappa\cos\theta$, in terms of which the drive Hamiltonian becomes
 \begin{multline}
@@
 where $m\in\set{+1,0,-1}$ indexes the laser pointing along $\v v_m$; and $s_{0,j}\equiv s_{\z,j}$ for shorthand.
-In the far-detuned limit $\Delta\gg\abs{\Omega_m}$, a second-order perturbative treatment of ${^3}\t{P}_0$ excitations ($\ket\e$) yields an effective drive Hamiltonian that only addresses ground-state nuclear spins.
+In the far-detuned limit $\abs{\Delta}\gg\abs{\Omega_m}$, a second-order perturbative treatment of electronic excitations ($\ket\e$) yields an effective drive Hamiltonian that only addresses ground-state nuclear spins.
 After a gauge transformation $c_{j\mu}^\dag\to e^{\i\phi\mu j} c_{j\mu}^\dag$ (equivalently $s_{mj}\to e^{\i m\phi j}s_{mj}$), we can then make the simplifying assumption (or requirement) that all $\Omega_m$ are real to write
@@
 
-Eqs.~\eqref{eq:drive_all} and \eqref{eq:drive_single} are key results of this section.
+Eqs.~\eqref{eq:drive_all} and \eqref{eq:drive_single} are the main technical results of this section.
 There are three important observations to make about these results.
-First, the fact that $H_{\t{3LD}}$ acts identically on all spins means we can freely replace the site index $j$ with a momentum index $q$ (as can be verified by substituting $c_{j\mu}=\frac1{\sqrt{L}}\sum_k e^{-\i q\c j} c_{q\mu}$), which is important to ensure that this drive addresses the same spin degrees of freedom as the final Hamiltonians in Section \ref{sec:SOC}.
+First, the fact that $H_{\t{3LD}}$ acts identically on all spins means we can freely replace the site index $j$ with a momentum index $q$ (as can be verified by substituting $c_{j\mu}=\frac1{\sqrt{L}}\sum_k e^{-\i q\c j} c_{q\mu}$), which is important to ensure that this drive addresses the same spin degrees of freedom as the spin Hamiltonian $H_{\t{spin}}$ in Eq.~\eqref{eq:H_spin}.
 Second, each of $\tilde\Omega_0,\tilde\Omega_+,\tilde\Omega_-$ can be tuned independently by changing the amplitudes of the driving lasers; some particular Hamiltonians for specific values of these amplitudes are shown in Table \ref{tab:drives}.
@@
 The capability to perform arbitrary spatial rotations, together with the capability to measure the number of atoms with spin projection $\mu$ onto a fixed quantization axis, $\bk{S_{\mu\mu}}$ (where $S_{\mu\nu}=\sum_js_{\mu\nu j}$), implies the capability to reconstruct all components of the mean collective spin matrix $\bk{\v S}=\sum_{\mu\nu}\bk{S_{\mu\nu}}\op{\mu}{\nu}$ via spin qudit tomography \cite{newton1968measurability, perlin2020qudit}.
+We leave open the question of whether and how a suitable choice of time-dependent amplitudes $\tilde\Omega_m$ in Eq.~\eqref{eq:drive_single} enables generating arbitrary SU($n$) rotations.
 
-Finally, we comment on initial states that these external controls allow us to prepare.
+Finally, we comment on the preparation of initial states.
+Initial states are nominally prepared in the ``lab gauge'', and must be transformed according to the gauge transformation $c_{j\mu}^\dag\to e^{\i\phi\mu j} c_{j\mu}^\dag$ prior to evolution under the ``gauged'' spin Hamiltonian $H_{\t{spin}}$ in Eq.~\eqref{eq:H_spin}.
+We assume the capability to prepare an initial state in which all spins are maximally polarized along $\v v$, i.e.~$\ket{\v v}^{\otimes N}=\ket{s}^{\otimes N}$, which is unaffected by the gauge transformation (up to a global phase).
+The three-laser then allows us to rotate this state into one that is polarized along any spatial axis (in the gauged frame).
+In addition, when $n>2$ the three-laser drive allows us to prepare product states with nontrivial intra-spin correlations.
+For example, by choosing $H_{\t{drive}}^{\t{single}}\propto s_\x^2$ (see Table \ref{tab:drives}), we can transform the polarized state $\ket{s}^{\otimes N}$ into $\p{\ket{+s}+\ket{-s}}^{\otimes N}$, where every spin is in a coherent superposition of states pointing in opposite directions along $\v v$.
+When $n>2$, this state has a vanishing mean spin vector, $\bk{S_\x}=\bk{S_\y}=\bk{S_\z}=0$ (where $S_\alpha = \sum_j s_{\alpha j}$), but $\bk{S_\z^2}=Ns^2$.
 
-\red{comment about initial states (``intra-spin correlations'')}
-
-\newpage
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{A simple, realistic model}
-\label{sec:full_model}
+\section{A simple model}
+\label{sec:simple_model}
 
@@
 
+\onecolumngrid
+\appendix
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Schwinger boson equations of motion}
+\label{sec:bosons}
+
+Here we decompose a quadratic spin Hamiltonian into Schwinger bosons, and derive the equations of motion for the resulting boson operators.
+We begin with a general spin Hamiltonian of the form
+\begin{align}
+  H = \sum_{\substack{\mu,\nu,\rho,\sigma\\j<k}}
+  h^{\mu\nu j}_{\rho\sigma k} s_{\mu\nu j} s_{\rho\sigma k}
+  + \sum_{\mu,\nu,j} \epsilon_{\mu\nu j} s_{\mu\nu j},
+  \label{eq:spin}
+\end{align}
+where $\mu,\nu$ index orthogonal states of an $n$-level spin; $j,k$ index one of $N$ spins; $h^{\mu\nu j}_{\rho\sigma k}$ and $\epsilon_{\mu\nu j}$ are scalars; and $s_{\mu\nu j}=\op{\mu}{\nu}_j$ is a transition operator for spin $j$.
+Strictly speaking, Eq.~\eqref{eq:spin} only defines the couplings $h^{\mu\nu j}_{\rho\sigma k}$ for $j<k$, so we enforce $h^{\mu\nu k}_{\rho\sigma j}=h^{\mu\nu j}_{\rho\sigma k}$ and $h^{\mu\nu j}_{\rho\sigma j}=0$ for completion.
+Decomposing spin operators into Schwinger bosons as $s_{\mu\nu j}=b_{\mu j}^\dag b_{\nu j}$, where $b_{\nu j}$ a annihilates a boson of type $\nu$ on site $j$, we can write this Hamiltonian as
+\begin{align}
+  H = \sum_{\substack{\mu,\nu,\rho,\sigma\\j<k}}
+  h^{\mu\nu j}_{\rho\sigma k}
+  b_{\mu j}^\dag b_{\nu j} b_{\rho k}^\dag b_{\sigma k}
+  + \sum_{\mu,\nu,j} \epsilon_{\mu\nu j} b_{\mu j}^\dag b_{\nu j}.
+\end{align}
+The Heisenberg equations of motion for the boson operators are then
+\begin{align}
+  \i \d_t b_{\alpha\ell} = \sp{b_{\alpha\ell}, H}
+  &= \sum_{\substack{\mu,\nu,\rho,\sigma\\j<k}}
+  h^{\mu\nu j}_{\rho\sigma k}
+  \sp{b_{\alpha\ell}, b_{\mu j}^\dag b_{\nu j} b_{\rho k}^\dag b_{\sigma k}}
+  + \sum_{\mu,\nu,j} \epsilon_{\mu\nu j}
+  \sp{b_{\alpha\ell}, b_{\mu j}^\dag b_{\nu j}} \\
+  &= \sum_{\mu,\nu,\rho,\sigma,k} h^{\mu\nu\ell}_{\rho\sigma k}
+  \sp{b_{\alpha\ell}, b_{\mu\ell}^\dag b_{\nu\ell}}
+  b_{\rho k}^\dag b_{\sigma k}
+  + \sum_{\mu,\nu} \epsilon_{\mu\nu\ell}
+  \sp{b_{\alpha\ell}, b_{\mu\ell}^\dag b_{\nu\ell}} \\
+  &= \sum_{\mu,\nu} \p{\sum_{\rho,\sigma,k}
+    h^{\mu\nu\ell}_{\rho\sigma k} b_{\rho k}^\dag b_{\sigma k}
+    + \epsilon_{\mu\nu\ell}}
+  \sp{b_{\alpha\ell}, b_{\mu\ell}^\dag b_{\nu\ell}}
+\end{align}
+where
+\begin{align}
+  \sp{b_{\alpha\ell}, b_{\mu\ell}^\dag b_{\nu\ell}}
+  = \delta_{\alpha\mu} \delta_{\alpha\nu} b_{\alpha\ell}
+  + \delta_{\alpha\mu} \p{1-\delta_{\alpha\nu}} b_{\nu\ell}
+  = \delta_{\alpha\mu} b_{\nu\ell},
+\end{align}
+so
+\begin{align}
+  \i \d_t b_{\alpha\ell}
+  = \sum_\nu \p{\sum_{\rho,\sigma,k}
+    h^{\alpha\nu\ell}_{\rho\sigma k} b_{\rho k}^\dag b_{\sigma k}
+    + \epsilon_{\alpha\nu\ell}} b_{\nu\ell}.
+\end{align}
+In the case of uniform SU($n$)-symmetric interactions of the form $\frac{h}{2}\v S\c\v S$ and a diagonal external field, we have
+\begin{align}
+  h^{\alpha\nu\ell}_{\rho\sigma k}
+  = h \times \delta_{\alpha\sigma} \delta_{\nu\rho},
+  &&
+  \epsilon_{\alpha\nu\ell}
+  = \epsilon_{\alpha\ell} \times \delta_{\alpha\nu}
+\end{align}
+so
+\begin{align}
+  \i \d_t b_{\alpha\ell}
+  = h \sum_{\nu,k} b_{\nu k}^\dag b_{\alpha k} b_{\nu\ell}
+  + \epsilon_{\alpha\ell} b_{\alpha\ell}.
+\end{align}
+
 \end{document}

### 2021-03-24 sun_phases/spin_model_paper/main.tex
moving spin model paper

--- 2021-03-22 sun_phases/sun_phases.tex
@@
 \newcommand{\e}{\text{e}}
+\renewcommand{\SS}{\mathbb{S}}
+\renewcommand{\O}{\mathcal{O}}
+\newcommand{\I}{\mathcal{I}}
 \newcommand{\su}{\mathfrak{su}}
-\renewcommand{\SS}{\mathbb{S}}
+\newcommand{\MFT}{\text{MFT}}
+
+\renewcommand{\ss}{\bar{\v s}\c\bar{\v s}}
 
@@
 
-\red{[finish intro, summarize paper]}
-% (un)interested reader can ...
+In this work, we ... \red{[finish intro, summarize paper]}
+% the (un)interested reader can ...
 % main results in/below Eqs. ...
@@
 \centering
-\includegraphics[width=\linewidth]{sketches/spin_model.pdf}
+\includegraphics[width=\linewidth]{spin_model.pdf}
 \caption{
@@
 \centering
-\includegraphics[width=\linewidth]{sketches/qudit_rep.pdf}
+\includegraphics[width=\linewidth]{qudit_rep.pdf}
 \caption{
@@
 \centering
-\includegraphics{sketches/soc_panels.pdf}
+\includegraphics{soc_panels.pdf}
 \caption{
@@
 However, $H_{\t{int}}^{(\phi)}$ cannot be exactly reduced to a spin Hamiltonian that addresses the same spin degrees of freedom as $H_{\t{kin}}^{(\phi)}$.
-Nonetheless, in the weak SOC limit $s\phi\ll1$ the effect of $H_{\t{int}}^{(\phi)}$ may be approximated by $H_{\t{int}}^{(0)}=H_{\t{int}}$, which has previously been benchmarked for SU(2)-symmetric interactions \cite{he2019engineering, smale2019observation}.
+Nonetheless, for sufficiently weak SOC (small $\phi$) the effect of $H_{\t{int}}^{(\phi)}$ may be approximated by $H_{\t{int}}^{(0)}=H_{\t{int}}$, which has previously been benchmarked for SU(2)-symmetric interactions \cite{he2019engineering, smale2019observation}.
 To ensure that $H_{\t{kin}}^{(\phi)}$ does not become trivial as $\phi\to0$, we can simultaneously take $J\to\infty$ while keeping $J\phi$ constant.
@@
 Altogether, the multi-laser drive Hamiltonian can be written in the form
-\begin{multline}
+\begin{align}
   H_{\t{drive}}^{\t{full}}
   = \sum_{j,\v v,\sigma} \Omega_{\v v\sigma}
-  \p{e^{-\i\kappa\v v\c\v\ell j} s_{\v v\sigma j}\otimes\op{\e}{\g}_j + \t{h.c.}} \\
-  + \Delta \sum_j \1_j \otimes \op{\e}_j,
+  \p{e^{-\i\kappa\v v\c\v\ell j} s_{\v v\sigma j}\otimes\op{\e}{\g}_j + \t{h.c.}}
+  + \Delta N_\e,
   \label{eq:drive_full}
-\end{multline}
-where $\sigma\in\set{+1,-1}$ indexes a polarization of light; $s_{\v v\sigma j}$ is a $n$-level spin-raising ($\sigma=+1$) or spin-lowering ($\sigma=-1$) operator along axis $\v v$ for atom $j$ (clarified below); $\ket\g_j$ and $\ket\e_j$ respectively denote the ground (${^1}\t{S}_0$) and excited (${^3}\t{P}_0$) electronic states of an atom on site $j$; $\1_j\equiv\sum_\mu s_{\mu\mu j}$ is an identity operator for the spin of atom $j$; and we work in a rotating frame in which the energy difference between ground and excited electronic states is $\Delta$.
+\end{align}
+where $\sigma\in\set{+1,-1}$ indexes a polarization of light; $s_{\v v\sigma j}$ is a $n$-level spin-raising ($\sigma=+1$) or spin-lowering ($\sigma=-1$) operator along axis $\v v$ for atom $j$ (clarified below); $\ket\g_j$ and $\ket\e_j$ respectively denote the ground (${^1}\t{S}_0$) and excited (${^3}\t{P}_0$) electronic states of an atom on site $j$; and $N_\e=\1\otimes\sum_j\op\e_j$ is the number of excited atoms (with $\1$ the identity operator on all spin degrees of freedom).
 
@@
 \centering
-\includegraphics[width=\linewidth]{sketches/3LD.pdf}
+\includegraphics[width=\linewidth]{3LD.pdf}
 \caption{
@@
 Choosing $\v v_+$ as the spin quantization axis, we can define new drive amplitudes $\Omega_0\equiv-(\Omega_{\v v_0,+}+\Omega_{\v v_0,-})$ and $\Omega_\pm\equiv\pm\Omega_{\v v_\pm,+}$, as well as a SOC angle $\phi\equiv\kappa\v v_+\c\v\ell=\kappa\cos\theta$, in terms of which the drive Hamiltonian becomes
-\begin{multline}
+\begin{align}
   H_{\t{3LD}}^{\t{full}}
   = \sum_{j,m} \Omega_m
-  \p{e^{-\i m\phi j} s_{mj} \otimes\op{\e}{\g}_j + \t{h.c.}} \\
-  + \Delta \sum_j \1_j \otimes \op{\e}_j,
-\end{multline}
+  \p{e^{-\i m\phi j} s_{mj} \otimes\op{\e}{\g}_j + \t{h.c.}}
+  + \Delta N_\e,
+\end{align}
 where $m\in\set{+1,0,-1}$ indexes the laser pointing along $\v v_m$; and $s_{0,j}\equiv s_{\z,j}$ for shorthand.
@@
 Finally, we comment on the preparation of initial states.
-Initial states are nominally prepared in the ``lab gauge'', and must be transformed according to the gauge transformation $c_{j\mu}^\dag\to e^{\i\phi\mu j} c_{j\mu}^\dag$ prior to evolution under the ``gauged'' spin Hamiltonian $H_{\t{spin}}$ in Eq.~\eqref{eq:H_spin}.
+Initial states are nominally prepared in the ``lab frame'', and must be transformed according to the gauge transformation $c_{q\mu}^\dag\to c_{q-\mu\phi,\mu}^\dag$ prior to evolution under the ``gauged frame'' spin Hamiltonian $H_{\t{spin}}$ in Eq.~\eqref{eq:H_spin} or the three-laser drive $H_{\t{3LD}}$ in Eq.~\eqref{eq:drive_all}.
 We assume the capability to prepare an initial state in which all spins are maximally polarized along $\v v$, i.e.~$\ket{\v v}^{\otimes N}=\ket{s}^{\otimes N}$, which is unaffected by the gauge transformation (up to a global phase).
@@
 In addition, when $n>2$ the three-laser drive allows us to prepare product states with nontrivial intra-spin correlations.
-For example, by choosing $H_{\t{drive}}^{\t{single}}\propto s_\x^2$ (see Table \ref{tab:drives}), we can transform the polarized state $\ket{s}^{\otimes N}$ into $\p{\ket{+s}+\ket{-s}}^{\otimes N}$, where every spin is in a coherent superposition of states pointing in opposite directions along $\v v$.
+For example, by choosing $H_{\t{drive}}^{\t{single}}\propto s_\x^2$ (see Table \ref{tab:drives}), we can transform the polarized state $\ket{s}^{\otimes N}$ into $\p{\ket{s}+\ket{-s}}^{\otimes N}$, where every spin is in a coherent superposition of states pointing in opposite directions along $\v v$.
 When $n>2$, this state has a vanishing mean spin vector, $\bk{S_\x}=\bk{S_\y}=\bk{S_\z}=0$ (where $S_\alpha = \sum_j s_{\alpha j}$), but $\bk{S_\z^2}=Ns^2$.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{A simple model}
-\label{sec:simple_model}
+\section{Mean-field theory and dynamical phases}
+\label{sec:mean_field}
 
-% use/simulate sinusoidal dispersion
+We now return to the spin Hamiltonian $H_{\t{spin}}$ in Eq.~\eqref{eq:H_spin}, and henceforth work exclusively in the ``gauged frame'' of this Hamiltonian and the three-laser drive $H_{\t{3LD}}$ in Eq.~\eqref{eq:drive_all}.
+At a fixed atom number $N$, the spin Hamiltonian $H_{\t{spin}}$ essentially has one free parameter: the ratio $J\phi/U$ that determines the relative strength of the single-particle and interaction terms.
+On a high level, for an initial $N$-fold tensor product state $\ket\psi^{\otimes N}$ one should expect distinct dynamical behaviors when $J\phi/U\ll1$, in which case spin-locking through strong $\v S\c\v S$ interactions restricts dynamics to the permutationally symmetric manifold, as opposed to $J\phi/U\gg1$, in which case individual spins should dephase due to the inhomogeneous magnetic field.
+To investigate these behaviors quantitatively at the mean-field level, we first decompose individual spin operators into Schwinger bosons as $s_{\mu\nu q} = b_{\mu q}^\dag b_{\nu q}$, such that the spin Hamiltonian $H_{\t{spin}}$ in Eq.~\eqref{eq:H_spin} becomes
+\begin{align}
+  H_{\t{boson}} = -\f{U}{2N} \sum_{p,q,\mu,\nu}
+  b_{\mu p}^\dag b_{\nu p} b_{\nu q}^\dag b_{\mu q}
+  + J\phi\sum_{q,\mu} \mu\, b_{\mu q}^\dag b_{\mu q}.
+\end{align}
+The Heisenberg equations of motion for the Schwinger boson operators are (see Appendix \ref{sec:bosons})
+\begin{align}
+  \i \d_t b_{\mu q}
+  = -\f{U}{N} \sum_{\nu,p} b_{\nu p}^\dag b_{\mu p} b_{\nu q}
+  + \mu\, b_{\mu q}.
+\end{align}
+Our mean-field theory (MFT) then treats all boson operators in these equations of motion as complex numbers, $b_{\mu q}\to\bk{b_{\mu q}}_\MFT$, with the initial value $\bk{b_{\mu q}\p{t=0}}_\MFT$ equal to the initial amplitude of spin $q$ in state $\mu$.
+Specifically, we will consider initial $N$-fold product states $\ket\psi^{\otimes N}$ with $\ket\psi = \sum_\mu \psi_\mu \ket\mu$, for which we set all $\bk{b_{\mu q}\p{t=0}}_\MFT = \psi_\mu$.
+For pure initial product states, this mean-field treatment of the boson operators $b_{\mu q}$ is mathematically equivalent to a mean-field treatment of the spin operators $s_{\mu\nu q}$, but reduces the number of variables to keep track of by a factor of $\sim n$.
+
+\vspace{1cm}
+
+\red{[Text that will depend on the figures]}
+
+\vspace{1cm}
+
+To identify dynamical phases, we examine long-time-averaged observables of the form
+\begin{align}
+  \bbk{\O}_\MFT = \lim_{T\to\infty} \f1T \int_0^T \dd t\, \bk{\O\p{t}}_\MFT,
+\end{align}
+where $\bk{\O\p{t}}_\MFT$ is the mean-field value of $\O$ at time $t$.
+Specifically, we consider the observables
+\begin{align}
+  \bar\sigma_\x \equiv \f{S_\x}{\norm{S_\x}},
+  &&
+  \ss = \f{\v S}{N}\c\f{\v S}{N},
+\end{align}
+where $S_\x=\sum_q s_{\x,q}$ is a collective spin-$x$ operator with operator norm $\norm{S_\x}=Ns$, and $\bar{\v s}\equiv\frac1N\sum_q\v s_q$ is the average spin matrix (related to the average single-particle density operator $\bar\rho$ by $\bar{\v s} = \bar\rho^\dag$).
+By design, these observables to have norm $\norm{\bar\sigma_\x}=\norm{\ss}=1$ independent of the system size $N$ or spin dimension $n$.
+
+\vspace{1cm}
+
+\red{[Discussion of figures]}
+
+\vspace{1cm}
+
+\begin{figure}
+\centering
+\includegraphics{mean_x_ss.pdf}
+\caption{
+Long-time averages of $\bar\sigma_\x$ and $\ss$ for different spin dimensions $n$ (indicated in the legend) as determined by mean-field simulations of $N=100$ spins up to a maximal time $T=10^4/U$.
+}
+\label{fig:mean_x_ss}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{crit_fields.pdf}
+\caption{
+The critical value of $\p{J\phi/U}_{\t{crit}}$ as determined by mean-field simulations of $N=100$ spins, and a single-parameter fit to $\p{J\phi/U}_{\t{crit}}=\p{n/2}^\alpha$, which finds $\alpha=-0.333(4)$.
+The fit agrees with all mean-field results to within an uncertainty determined by the resolution of $J\phi/U$ in mean-field simulations.
+}
+\label{fig:crit_fields}
+\end{figure}
+
 % initial state in |X>
@@
 
-\bibliography{\jobname.bib}
-
+\bibliography{main.bib}
 \onecolumngrid
@@
   + \sum_{\mu,\nu,j} \epsilon_{\mu\nu j} s_{\mu\nu j},
-  \label{eq:spin}
+  \label{eq:quadratic_spin}
 \end{align}
 where $\mu,\nu$ index orthogonal states of an $n$-level spin; $j,k$ index one of $N$ spins; $h^{\mu\nu j}_{\rho\sigma k}$ and $\epsilon_{\mu\nu j}$ are scalars; and $s_{\mu\nu j}=\op{\mu}{\nu}_j$ is a transition operator for spin $j$.
-Strictly speaking, Eq.~\eqref{eq:spin} only defines the couplings $h^{\mu\nu j}_{\rho\sigma k}$ for $j<k$, so we enforce $h^{\mu\nu k}_{\rho\sigma j}=h^{\mu\nu j}_{\rho\sigma k}$ and $h^{\mu\nu j}_{\rho\sigma j}=0$ for completion.
+Strictly speaking, Eq.~\eqref{eq:quadratic_spin} only defines the couplings $h^{\mu\nu j}_{\rho\sigma k}$ for $j<k$, so we enforce $h^{\mu\nu k}_{\rho\sigma j}=h^{\mu\nu j}_{\rho\sigma k}$ and $h^{\mu\nu j}_{\rho\sigma j}=0$ for completion.
 Decomposing spin operators into Schwinger bosons as $s_{\mu\nu j}=b_{\mu j}^\dag b_{\nu j}$, where $b_{\nu j}$ a annihilates a boson of type $\nu$ on site $j$, we can write this Hamiltonian as

### 2021-03-24 sun_phases/spin_model_paper/figures/dicke_methods.py

@@ -1,0 +1,1 @@
+../../scripts/dicke_methods.py

### 2021-03-24 sun_phases/spin_model_paper/figures/make_plots.py

@@ -235,15 +235,26 @@ def plot_sphere(vectors, points = None, color = "#d62728", width = 2):
     sphere.point_color = [ color ]
 
     sphere.render()
+    plt.gca().set_zlim(-.55,.55)
     return sphere
 
 sphere = plot_sphere([0,1,0])
 plt.savefig("bloch_x.pdf")
 
+def vec_xy(angle = 0):
+    xhat = np.array([1,0,0])
+    yhat = np.array([0,1,0])
+    return np.cos(angle) * yhat - np.sin(angle) * xhat
+
+angle = 2*np.pi * 2/3
+points = 10
+points = [ vec_xy(part) for part in np.linspace(0,angle,points) ]
+sphere = plot_sphere(vec_xy(angle), points)
+plt.savefig("bloch_xy.pdf")
+
 import scipy.stats
 from dicke_methods import coherent_spin_state, spin_op_z_dicke, plot_dicke_state
 
-grid_size = 501
 kwargs = dict( single_sphere = True, shade = False, grid_size = 501 )
 
 qubit_state = coherent_spin_state("+X", 1)

### 2021-03-24 sun_phases/spin_model_paper/figures/special_functions.py

@@ -1,0 +1,1 @@
+../../scripts/special_functions.py

### 2021-03-30 sun_phases/spin_model_paper/main.tex
updating multilevel spin model paper

@@
 \newcommand{\su}{\mathfrak{su}}
-\newcommand{\MFT}{\text{MFT}}
+\newcommand{\MF}{\text{MF}}
 
 \renewcommand{\ss}{\bar{\v s}\c\bar{\v s}}
+\DeclareMathOperator{\diag}{diag}
+
+%%% for in-line enumerated lists
+\usepackage[inline]{enumitem}
+\setlist[enumerate,1]{label={(\roman*)}} % default in-line numbering
+\setlist{nolistsep} % more compact spacing between environments
+
+% use alphanumeric tags for footnotes
+\renewcommand*{\thefootnote}{\alph{footnote}}
 
@@
 \newcommand{\red}[1]{{\color{red} #1}}
-
-% use alphanumeric tags for footnotes
-\renewcommand*{\thefootnote}{\alph{footnote}}
 
@@
 \begin{align}
-  \v A \c \v B \equiv \Tr\p{\v A^\dag\v B}
+  \v A \c \v B \equiv \Tr_{\t{aux}}\p{\v A^\dag\v B}
   = \sum_{\mu,\nu} A_{\mu\nu}^\dag B_{\mu\nu},
 \end{align}
-where the trace is performed only over the auxiliary matrix degrees of freedom.
+where the trace is performed over the auxiliary matrix degrees of freedom.
 These definitions allow us to write the spin Hamiltonian in Eq.~\eqref{eq:H_int_start} as
@@
 The dimension $n$ is reflected in $Q_\rho$ by the fact that it can be decomposed into spherical harmonics $Y_{\ell m}$ with degree $\ell<n$ (that is, $Q_\rho$ is ``low resolution'' or ``band-limited'' at degree $n$) \cite{perlin2020qudit}, as well as in the normalization $\int_{\v v\in\SS^2}\dd^2\v v\,Q_\rho\p{\v v}=\frac{4\pi}{d}$.
+Closely related spherical representations of multilevel spin operators are discussed in Ref.~\cite{li2013weylwignermoyal}.
 
@@
 The Hilbert space of a 10-level (spin-$\frac92$) nucleus of a ${}^{87}$Sr atom, for example, can be thought of as the Dicke manifold of 9 ``excess'' proton and neutron spins.
-This representation of a $n$-level quantum system is a useful picture to keep in mind for angular-momentum-like degrees of freedom, and will be used freely throughout this work.
 
@@
 \begin{align}
-  \left. H_{\t{kin}}^{(\phi)} \right|_{d=2}
+  \left. H_{\t{kin}}^{(\phi)} \right|_{n=2}
   = 2\sin\p{\phi/2} J \sum_q \sin\p{q} s_{\z,q}.
@@
 In the far-detuned limit $\abs{\Delta}\gg\abs{\Omega_m}$, a second-order perturbative treatment of electronic excitations ($\ket\e$) yields an effective drive Hamiltonian that only addresses ground-state nuclear spins.
-After a gauge transformation $c_{j\mu}^\dag\to e^{\i\phi\mu j} c_{j\mu}^\dag$ (equivalently $s_{mj}\to e^{\i m\phi j}s_{mj}$), we can then make the simplifying assumption (or requirement) that all $\Omega_m$ are real to write
+After a gauge transformation $c_{j\mu}^\dag\to e^{\i\phi\mu j} c_{j\mu}^\dag$ (equivalently $s_{mj}\to e^{\i m\phi j}s_{mj}$), we can then make the simplifying assumption that all $\Omega_m$ are real to write
 \begin{align}
@@
 Second, each of $\tilde\Omega_0,\tilde\Omega_+,\tilde\Omega_-$ can be tuned independently by changing the amplitudes of the driving lasers; some particular Hamiltonians for specific values of these amplitudes are shown in Table \ref{tab:drives}.
-Third, due to the appearance of mutually commuting pairs of Hamiltonians in Table \ref{tab:drives}, specifically $-s_\alpha^2$ and $\pm s_\alpha+s_\alpha^2$ for $\alpha\in\set{\z,\x}$, the three-laser drive admits pulse sequences that can implement arbitrary SU(2) (spatial) rotations of the form $e^{-\i\chi\vec n\c\vec s}$, where $\chi$ is a rotation angle, $\vec n$ is a rotation axis, and $\vec s\equiv(s_\x,s_\y,s_\z)$.
+Third, due to the appearance of mutually commuting pairs of Hamiltonians in Table \ref{tab:drives}, specifically $-s_\alpha^2$ and $\pm s_\alpha+s_\alpha^2$ for $\alpha\in\set{\z,\x}$, the three-laser drive admits pulse sequences that exactly implement arbitrary SU(2) (spatial) rotations of the form $e^{-\i\chi\vec n\c\vec s}$, where $\chi$ is a rotation angle, $\vec n$ is a rotation axis, and $\vec s\equiv(s_\x,s_\y,s_\z)$.
 The capability to perform arbitrary spatial rotations, together with the capability to measure the number of atoms with spin projection $\mu$ onto a fixed quantization axis, $\bk{S_{\mu\mu}}$ (where $S_{\mu\nu}=\sum_js_{\mu\nu j}$), implies the capability to reconstruct all components of the mean collective spin matrix $\bk{\v S}=\sum_{\mu\nu}\bk{S_{\mu\nu}}\op{\mu}{\nu}$ via spin qudit tomography \cite{newton1968measurability, perlin2020qudit}.
-We leave open the question of whether and how a suitable choice of time-dependent amplitudes $\tilde\Omega_m$ in Eq.~\eqref{eq:drive_single} enables generating arbitrary SU($n$) rotations.
+Moreover, we expect that advanced quantum optimal control techniques (similar to those of Refs.~\cite{anderson2015accurate, lucarelli2018quantum}) can be used to implement arbitrary SU($n$) rotations by designing suitable time-dependent drive amplitudes $\Omega_m$.
 
@@
 
-We now return to the spin Hamiltonian $H_{\t{spin}}$ in Eq.~\eqref{eq:H_spin}, and henceforth work exclusively in the ``gauged frame'' of this Hamiltonian and the three-laser drive $H_{\t{3LD}}$ in Eq.~\eqref{eq:drive_all}.
-At a fixed atom number $N$, the spin Hamiltonian $H_{\t{spin}}$ essentially has one free parameter: the ratio $J\phi/U$ that determines the relative strength of the single-particle and interaction terms.
-On a high level, for an initial $N$-fold tensor product state $\ket\psi^{\otimes N}$ one should expect distinct dynamical behaviors when $J\phi/U\ll1$, in which case spin-locking through strong $\v S\c\v S$ interactions restricts dynamics to the permutationally symmetric manifold, as opposed to $J\phi/U\gg1$, in which case individual spins should dephase due to the inhomogeneous magnetic field.
-To investigate these behaviors quantitatively at the mean-field level, we first decompose individual spin operators into Schwinger bosons as $s_{\mu\nu q} = b_{\mu q}^\dag b_{\nu q}$, such that the spin Hamiltonian $H_{\t{spin}}$ in Eq.~\eqref{eq:H_spin} becomes
+We now return to the spin Hamiltonian $H_{\t{spin}}$ in Eq.~\eqref{eq:H_spin}, which is valid in the weak-SOC limit $s\phi\ll1$, and henceforth work exclusively in the ``gauged frame'' of this $H_{\t{spin}}$ and the three-laser drive $H_{\t{3LD}}$ in Eq.~\eqref{eq:drive_all}.
+At the mean-field (MF) level, the spin Hamiltonian becomes
 \begin{align}
-  H_{\t{boson}} = -\f{U}{2N} \sum_{p,q,\mu,\nu}
+  H_\MF = \sum_q\sp{-U\bk{\bar{\v s}}\c\v s_q
+    + J\phi \sum_q \sin\p{q} s_{\z,q}}.
+  \label{eq:H_MF}
+\end{align}
+where $\bar{\v s}\equiv\frac1N\sum_q\v s_q$ is the average spin matrix.
+Fixing the atom number $N$, the spin Hamiltonian has one free parameter: the reduced field $J\phi/U$ that determines the relative strength of the single-particle and interaction terms.
+For an initial $N$-fold tensor product state $\ket\psi^{\otimes N}$ one should expect distinct dynamical behaviors when $J\phi/U\ll1$, in which case strong spin-locking interactions should give rise to a long-range ordered phase, as opposed to $J\phi/U\gg1$, in which case long-range order should be destroyed by the strong inhomogeneous magnetic field.
+
+To investigate these behaviors quantitatively, we examine long-time-averaged observables of the form
+\begin{align}
+  \bbk{\O}_\MF = \lim_{T\to\infty} \f1T \int_0^T \dd t\, \bk{\O\p{t}}_\MF,
+\end{align}
+where $\bk{\O\p{t}}_\MF$ is the mean-field value of observable $\O$ at time $t$ (see Appendices \ref{sec:MFT} and \ref{sec:bosons} for details about our mean-field simulations).
+Specifically, we consider long-time averages of the normalized transverse magnetization and interaction energy
+\begin{align}
+  \bar\sigma_\x \equiv \f{S_\x}{\norm{S_\x}},
+  &&
+  \ss = \f{\v S\c\v S}{N^2},
+\end{align}
+where $\norm{S_\x}=Ns$ is the operator norm of $S_\x$.
+By design, these observables to have norm $\norm{\bar\sigma_\x}=\norm{\ss}=1$ independent of the system size $N$ or spin dimension $n$.
+
+\begin{figure}
+\centering
+\includegraphics{mean_op_ss_X.pdf}
+\caption{
+Long-time averages of $\bar\sigma_\x$ and $\ss$ for different spin dimensions $n$ (indicated in the legend) as determined by mean-field simulations of $N=100$ spins initially in the $x$-polarized state $\ket\x^{\otimes N}$.
+Insets show same data with a rescaling $J\phi/U\to J\phi/U \times (n/2)^{1/3}$, $\bbk{\bar\sigma_x}_\MF\to\bbk{\bar\sigma_x}_\MF/\gamma\p{n/2}$, and $\bbk{\ss}_\MF\to\bbk{\ss}_\MF/\gamma\p{n}-1$, where $\gamma\p{k}$ is defined in Eq.~\eqref{eq:gamma}.
+}
+\label{fig:mean_op_ss_X}
+\end{figure}
+
+Figure \ref{fig:mean_op_ss_X} shows the long-time averages of the transverse magnetization $\bar\sigma_\x$ and interaction energy $\ss$ as computed by mean-field simulations of $N=100$ spins in an initial $x$-polarized state $\ket\x^{\otimes N}$, where
+\begin{align}
+  \ket\x \equiv e^{-\i\frac{\pi}{2}s_\y}\ket{s}
+  = \f1{2^s} \sum_\mu { 2s \choose s+\mu }^{1/2} \ket{\mu}.
+  \label{eq:state_x}
+\end{align}
+Here ${m \choose k}$ is a binomial coefficient.
+As expected, the spin model exhibits a mean-field dynamical phase transition between an ordered phase at small $J\phi/U$ and a disordered phase at large $J\phi/U$.
+The ordered phase has a non-zero transverse magnetization $\bbk{\sigma_\x}_\MF$ and a large interaction energy $\bbk{\ss}_\MF$, while the disordered phase has no transverse magnetization and a minimal interaction energy.
+Numerically, we find that the transition between these phases occurs at a critical field $\p{J\phi/U}_{\t{crit}}=\p{n/2}^{-\alpha}$ with $\alpha\approx1/3$ (see Figure \ref{fig:crit_fields_X}).
+
+\begin{figure}
+\centering
+\includegraphics{crit_fields_X.pdf}
+\caption{
+The critical value of $\p{J\phi/U}_{\t{crit}}$ as determined by mean-field simulations of $N=100$ spins initially in the $x$-polarized state $\ket\x^{\otimes N}$.
+A single-parameter fit to $\p{J\phi/U}_{\t{crit}}=\p{n/2}^{-\alpha}$ finds $\alpha=0.333(5)$, and $\alpha=1/3$ is consistent with all mean-field results to within an uncertainty determined by the resolution of $J\phi/U$ in mean-field simulations.
+}
+\label{fig:crit_fields_X}
+\end{figure}
+
+As shown in insets of Figure \ref{fig:mean_op_ss_X}, mean-field results for different spin dimensions $n$ collapse onto each other when normalizing the field $J\phi/U$ to its critical value, $\p{J\phi/U}\to\p{J\phi/U}\times\p{n/2}^{1/3}$, and rescaling $\bbk{\bar\sigma_x}_\MF\to\bbk{\bar\sigma_x}_\MF/\gamma\p{n/2}$ and $\bbk{\ss}_\MF\to\bbk{\ss}_\MF/\gamma\p{n}-1$, where
+\begin{align}
+  \gamma\p{k} \equiv \f{\Gamma\p{k-\frac12}}{\sqrt\pi\,\Gamma\p{k}}.
+  \label{eq:gamma}
+\end{align}
+The rescaling of transverse magnetization and interaction energy can be understood by considering their limiting behavior as $J\phi/U\to\infty$ or $J\phi/U\to0^+$.
+
+In the strong-field limit $J\phi/U\to\infty$, we can ignore interactions and treat spins as though they simply precess at different rates.
+The long-time-averaged transverse magnetization $\bbk{\bar\sigma_\x}_\MF$ then trivially vanishes as $J\phi/U\to\infty$.
+The interaction energy $\bk{\ss}_\MF = \bk{\bar{\v s}}_\MF \c \bk{\bar{\v s}}_\MF$, meanwhile, has contributions from
+\begin{enumerate*}
+\item the diagonal parts of the mean spin matrix $\bk{\bar{\v s}}_\MF$, which are conserved by inhomogeneous spin precession, and
+\item the off-diagonal parts of $\bk{\bar{\v s}}_\MF$, whose oscillations average to zero when evaluating the long-time average $\bbk{\ss}_\MF$.
+\end{enumerate*}
+Altogether, the interaction energy $\bbk{\ss}_\MF$ in the strong-field limit is determined by the time-independent diagonal part $\diag\bk{\bar{\v s}\p{t}} = \diag\op\x$, namely
+\begin{align}
+  \lim_{J\phi/U\to\infty} \bbk{\ss}_\MF
+  = \Tr\sp{\p{\diag\op\x}^2}
+  = \gamma\p{n} \approx \f1{\sqrt{2\pi s}}.
+\end{align}
+The same result can be obtained by considering the interaction energy of two counter-rotating spins, whose long-time average is essentially the time-averaged overlap of their phase-space distributions (see Figure \ref{fig:spin_dist}).
+
+In the weak-field limit $J\phi/U\to0^+$, the spin-locking $\v S\c\v S$ interactions of the spin Hamiltonian $H_{\t{spin}}$ effectively restrict dynamics to the permutationally-symmetric (PS) manifold.
+At second order in perturbation theory, the spin Hamiltonian thus becomes
+\begin{align}
+  H_{\t{spin}}^{\t{eff}}
+  = \f{\p{J\phi}^2}{\p{N-1}U}
+  \times \sp{S_\z^2 - N\sum_q s_{\z,q}^2}.
+\end{align}
+However, mean-field theory fails to capture inter-spin correlations that accumulate throughout evolution under $H_{\t{spin}}^{\t{eff}}$.
+For example, $\ss$ is constant (equal to 1) in the PS manifold, but $\lim_{J\phi/U\to0^+}\bbk{\ss}_\MF<1$ when $n>2$ in Figure \ref{fig:mean_op_ss_X}.
+Treating $H_{\t{spin}}^{\t{eff}}$ at the mean-field level would also yield a homogeneous Hamiltonian that conserves $\ss$, which is again inconsistent with the weak-field limits of $\bbk{\ss}_\MF$.
+This inconsistency between mean-field theories stems from the fact that applying perturbation theory does not commute with a mean-field factorization of spin operators.
+In order to reproduce the weak-field asymptotic values of $\bbk{\bar\sigma_\x}_\MF$ and $\bbk{\ss}_\MF$, therefore, we must instead consider the mean-field Hamiltonian $H_\MF$ in Eq.~\eqref{eq:H_MF}.
+
+\vspace{3cm}
+
+\bibliography{main.bib}
+\onecolumngrid
+\appendix
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Mean-field theory}
+\label{sec:MFT}
+
+Here we describe the mean-field theory used to simulate the spin Hamiltonian
+\begin{align}
+  H_{\t{spin}} = -\f{U}{2N}\v S\c\v S + J\phi \sum_q \sin\p{q} s_{\z,q}
+\end{align}
+in Eq.~\eqref{eq:H_spin} of the main text.
+We begin by decomposing individual spin operators into Schwinger bosons as $s_{\mu\nu q} = b_{\mu q}^\dag b_{\nu q}$, such that the spin Hamiltonian becomes
+\begin{align}
+  H_{\t{spin}} \to H_{\t{boson}}
+  = -\f{U}{2N} \sum_{p,q,\mu,\nu}
   b_{\mu p}^\dag b_{\nu p} b_{\nu q}^\dag b_{\mu q}
-  + J\phi\sum_{q,\mu} \mu\, b_{\mu q}^\dag b_{\mu q}.
+  + J\phi \sum_{q,\mu} \sin\p{q} \mu\, b_{\mu q}^\dag b_{\mu q}.
 \end{align}
@@
   = -\f{U}{N} \sum_{\nu,p} b_{\nu p}^\dag b_{\mu p} b_{\nu q}
-  + \mu\, b_{\mu q}.
+  + J\phi \sin\p{q} \mu\, b_{\mu q}.
 \end{align}
-Our mean-field theory (MFT) then treats all boson operators in these equations of motion as complex numbers, $b_{\mu q}\to\bk{b_{\mu q}}_\MFT$, with the initial value $\bk{b_{\mu q}\p{t=0}}_\MFT$ equal to the initial amplitude of spin $q$ in state $\mu$.
-Specifically, we will consider initial $N$-fold product states $\ket\psi^{\otimes N}$ with $\ket\psi = \sum_\mu \psi_\mu \ket\mu$, for which we set all $\bk{b_{\mu q}\p{t=0}}_\MFT = \psi_\mu$.
-For pure initial product states, this mean-field treatment of the boson operators $b_{\mu q}$ is mathematically equivalent to a mean-field treatment of the spin operators $s_{\mu\nu q}$, but reduces the number of variables to keep track of by a factor of $\sim n$.
-
-\vspace{1cm}
-
-\red{[Text that will depend on the figures]}
-
-\vspace{1cm}
-
-To identify dynamical phases, we examine long-time-averaged observables of the form
-\begin{align}
-  \bbk{\O}_\MFT = \lim_{T\to\infty} \f1T \int_0^T \dd t\, \bk{\O\p{t}}_\MFT,
-\end{align}
-where $\bk{\O\p{t}}_\MFT$ is the mean-field value of $\O$ at time $t$.
-Specifically, we consider the observables
-\begin{align}
-  \bar\sigma_\x \equiv \f{S_\x}{\norm{S_\x}},
-  &&
-  \ss = \f{\v S}{N}\c\f{\v S}{N},
-\end{align}
-where $S_\x=\sum_q s_{\x,q}$ is a collective spin-$x$ operator with operator norm $\norm{S_\x}=Ns$, and $\bar{\v s}\equiv\frac1N\sum_q\v s_q$ is the average spin matrix (related to the average single-particle density operator $\bar\rho$ by $\bar{\v s} = \bar\rho^\dag$).
-By design, these observables to have norm $\norm{\bar\sigma_\x}=\norm{\ss}=1$ independent of the system size $N$ or spin dimension $n$.
-
-\vspace{1cm}
-
-\red{[Discussion of figures]}
-
-\vspace{1cm}
-
-\begin{figure}
-\centering
-\includegraphics{mean_x_ss.pdf}
-\caption{
-Long-time averages of $\bar\sigma_\x$ and $\ss$ for different spin dimensions $n$ (indicated in the legend) as determined by mean-field simulations of $N=100$ spins up to a maximal time $T=10^4/U$.
-}
-\label{fig:mean_x_ss}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{crit_fields.pdf}
-\caption{
-The critical value of $\p{J\phi/U}_{\t{crit}}$ as determined by mean-field simulations of $N=100$ spins, and a single-parameter fit to $\p{J\phi/U}_{\t{crit}}=\p{n/2}^\alpha$, which finds $\alpha=-0.333(4)$.
-The fit agrees with all mean-field results to within an uncertainty determined by the resolution of $J\phi/U$ in mean-field simulations.
-}
-\label{fig:crit_fields}
-\end{figure}
-
-% initial state in |X>
-% repeat analysis with different initial state
-% -- look at class of initial states (e.g. back-to-back, tipping angle)
-
-\bibliography{main.bib}
-\onecolumngrid
-\appendix
+Our mean-field theory then treats all boson operators in these equations of motion as complex numbers, $b_{\mu q}\to\bk{b_{\mu q}}_\MF$, with the initial value $\bk{b_{\mu q}\p{t=0}}_\MF$ equal to the initial amplitude of spin $q$ in state $\mu$.
+Specifically, for an $N$-fold product state of the form $\ket\psi=\bigotimes_q\sum_\mu\psi_{\mu q}\ket{\mu}$ we set $\bk{b_{\mu q}\p{t=0}}_\MF = \psi_{\mu q}$.
+For pure initial product states, this mean-field treatment of the boson operators $b_{\mu q}$ is mathematically equivalent to a mean-field treatment of the spin operators $s_{\mu\nu q}$, as in Eq.~\eqref{eq:H_MF}, but reduces the number of variables to keep track of by a factor of $\sim n$.
 

### 2021-04-01 sun_phases/spin_model_paper/main.tex
updating multilevel spin model paper

@@
 \renewcommand{\SS}{\mathbb{S}}
+\newcommand{\A}{\mathcal{A}}
+\newcommand{\C}{\mathcal{C}}
+\newcommand{\I}{\mathcal{I}}
 \renewcommand{\O}{\mathcal{O}}
-\newcommand{\I}{\mathcal{I}}
+\renewcommand{\P}{\mathcal{P}}
+
 \newcommand{\su}{\mathfrak{su}}
 \newcommand{\MF}{\text{MF}}
-
 \renewcommand{\ss}{\bar{\v s}\c\bar{\v s}}
@@
   H_{\t{kin}}
-  &= -J\sum_q \cos\p{q} c_{q\mu}^\dag c_{q\mu}, \\
+  &= -J\sum_{q,\mu} \cos\p{q} c_{q\mu}^\dag c_{q\mu}, \\
   H_{\t{int}}
@@
 \end{align}
-where $N$ is the total number of atoms on the lattice, we define $U\equiv U_0\times N/L$ for convenience, $\delta_{k+p,\ell+q}=1$ if $k+p=\ell+q$ and $0$ otherwise (enforcing conservation of momentum), and we work in units with lattice spacing $a=1$.
+where $N$ is the total number of atoms on the lattice, we define $U\equiv U_0\times N/L$ for convenience, $\delta_{k+p,\ell+q}=1$ if $k+p=\ell+q$ and zero otherwise (enforcing conservation of momentum), and we work in units with lattice spacing $a=1$.
 
@@
 \begin{align}
-  H_{\t{kin}} &= -J\sum_q \cos\p{q} s_{\mu\mu q},
+  H_{\t{kin}} &= -J\sum_{q,\mu} \cos\p{q} s_{\mu\mu q},
   \label{H_kin_start} \\
@@
 \end{align}
-Throughout this work, we will assume that atomic modes are singly-occupied, e.g.~due to the initialization of a spin-polarized state with one atom per lattice site, in which multiple occupation of an atomic mode is forbidden by Pauli exclusion.
+Throughout this work, we will assume that atomic modes are singly-occupied, e.g.~due to the initialization of a spin-polarized state with one atom per lattice site, in which multiple occupation of an atomic mode is forbidden by fermionic statistics (Pauli exclusion).
 In this case, we can neglect the fermionic origins of $H_{\t{int}}$ and instead treat our system as a bona fide spin model.
+Note that the ``kinetic'' terms of this spin model ($H_{\t{kin}}$) are proportional to the identity operator, and can be neglected for the time being.
+The fermionic origin of these terms will return to play an important role when we consider spin-orbit coupling in Section \ref{sec:SOC}.
 
-To further simplify notation and write the interaction Hamiltonian in a form suggestively similar to SU(2) spin alignment, we now construct the operator-valued spin matrix
+To further simplify the interaction Hamiltonian $H_{\t{int}}$ and write it in a form suggestively similar to SU(2) spin alignment, we now construct the operator-valued spin matrix
 \begin{align}
@@
 \end{align}
-Here $\v S$ is a collective spin matrix, analogous to the collective spin vector $\vec S=\p{S_\x,S_\y,S_\z}$ in the case of SU(2) \cite{he2019engineering}.
+Here $\v S$ is a collective spin matrix, analogous to the collective spin vector $\vec S=\p{S_\x,S_\y,S_\z}$ in the case of SU(2) \cite{he2019engineering}, with $\v S\c\v S \simeq 2 \vec S\c\vec S = 2 \sum_{\alpha\in\set{\x,\y,\z}}S_\alpha^2$ when $n=2$ (with $\simeq$ denoting equality after modding out identity terms).
 
@@
 The operator $\v s_p\c\v s_q$ simply swaps the nuclear spin states of two atoms pinned to modes $p,q$.
-The term $\v s_p\c\v s_q$ thereby assigns a definite energy of $+1$ ($-1$) to a pair of spins that are symmetric (anti-symmetric) under exchange.
-In this sense, $\v s_p\c\v s_q$ is analogous to the enforcement of SU(2) spin alignment, which similarly assigns (distinct) definite energies to anti-symmetric spin-0 singlet $\ket{\up\dn}-\ket{\dn\up}$ and the symmetric spin-1 triplets $\set{\ket{\up\up},\ket{\dn\dn},\ket{\up\dn}+\ket{\dn\up}}$.
+The term $-\v s_p\c\v s_q$ thereby assigns a definite energy of $-1$ ($+1$) to a pair of spins that are symmetric (anti-symmetric) under exchange.
+In this sense, $\v s_p\c\v s_q$ is analogous to the enforcement of SU(2) spin alignment by ferromagnetic interactions, which similarly assigns (distinct) definite energies to anti-symmetric spin-0 singlet $\ket{\up\dn}-\ket{\dn\up}$ and the symmetric spin-1 triplets $\set{\ket{\up\up},\ket{\dn\dn},\ket{\up\dn}+\ket{\dn\up}}$.
 
@@
 In the general case of SU($n$), the PS manifold is similarly spanned by states $\ket{m_s,m_{s-1},\cdots,m_{-s}}$ with a definite number $m_\mu$ of spins in state $\mu$, and $\sum_\mu m_\mu=N$.
-The dimension of the PS manifold, or the ground-state degeneracy of $H_{\t{int}}$, is equal to the number of ways of assigning $N$ identical spins to $n$ distinct internal states, or ${N+n-1 \choose n-1} \sim N^{n-1}$.
+The dimension of the PS manifold is equal to the number of ways of assigning $N$ identical spins to $n$ distinct internal states, or ${N+n-1 \choose n-1} \sim N^{n-1}$.
+
+External fields or additional interactions that respect permutational symmetry can induce nontrivial dynamics within the PS manifold.
+Even if additional Hamiltonian terms break permutational symmetry, as long as the coupling to states outside of the PS manifold is small (compared to the energy gap $U$) these terms can be restricted to the PS manifold perturbatively (see Appendix \ref{sec:pert_theory}).
+Simulating dynamics within the PS manifold requires calculating matrix elements $\bk{\ell|\O|m}$ of spin operators $\O$ with respect to PS states $\ket\ell,\ket{m}$; we discuss this calculation in Appendix \ref{sec:PS_ops}.
 
@@
 Whenever possible, it is desirable to find geometric representations of abstract mathematical objects, such as the state of a quantum system.
-Geometric representations facilitate the development of understanding and intuitions that can be difficult to extract from algebraic expressions.
+Geometric representations facilitate the development of understanding and intuitions that can be difficult to extract from algebraic expressions alone.
 For this reason, the state of a two-level spin, or a qubit, is commonly represented by a point on (or within) the Bloch sphere.
@@
 The dimension $n$ is reflected in $Q_\rho$ by the fact that it can be decomposed into spherical harmonics $Y_{\ell m}$ with degree $\ell<n$ (that is, $Q_\rho$ is ``low resolution'' or ``band-limited'' at degree $n$) \cite{perlin2020qudit}, as well as in the normalization $\int_{\v v\in\SS^2}\dd^2\v v\,Q_\rho\p{\v v}=\frac{4\pi}{d}$.
-Closely related spherical representations of multilevel spin operators are discussed in Ref.~\cite{li2013weylwignermoyal}.
+Closely related spherical representations of multilevel spin states (and operators) are discussed in Ref.~\cite{li2013weylwignermoyal}.
 
@@
 We wish to study the dynamical behaviors of experimentally realizable SU($n$) spin models.
-In order to ensure validity of the $\v S\c\v S$ spin Hamiltonian in Eq.~\eqref{eq:H_int}, we consider initial preparation of a spin-polarized state\footnote{We remind the reader that validity of the spin model requires that each momentum mode is at most singly-occupied, which is guaranteed in a spin-polarized state due to Pauli blocking.}, which has the added benefit of being simple to prepare experimentally.
-However, a spin polarized state is an eigenstate of both the kinetic and interaction spin Hamiltonians in Section \ref{sec:spin_model}, so we need additional ingredients for nontrivial behavior.
+In order to ensure validity of the $\v S\c\v S$ spin Hamiltonian in Eq.~\eqref{eq:H_int}, we consider initial preparation of a spin-polarized state\footnote{We remind the reader that validity of the spin model requires that each momentum mode is at most singly-occupied, which is guaranteed for a spin-polarized state due to Pauli blocking.}, which has the added benefit of being simple to prepare experimentally.
+However, a spin polarized state is an eigenstate of the $\v S\c\v S$ spin Hamiltonian, so we need additional ingredients for nontrivial behavior.
 Here, we consider the addition of spin-orbit coupling (SOC) induced by external driving fields.
@@
 \end{align}
-is an $n$-level spin-$x$ operator for spin $q$, with which $e^{-\i\theta s_{\x,q}}$ rotates the state of spin $q$ about the $x$ axis by an angle $\theta$.
+is an $n$-level spin-$x$ operator for spin $q$, with which $e^{-\i\theta s_{\x,q}}$ rotates the state of spin $q$ about the $x$ axis by an angle $\theta$ (literally rotating the sphere in Figure \ref{fig:spin_dist}).
 Despite complicating details, the change from two-level to $n$-level spins thus results in a drive Hamiltonian with an identical physical interpretation: after moving into an appropriate ``gauged frame'', this drive simply rotates all spins about the $x$ axis at a rate $\Omega$.
@@
 In the ``lab frame'' ({\bf a},{\bf b}), kinetic energy is insensitive to spin, but the drive imparts a momentum kick $q\to q+\phi$ into any atom that absorbs a photon.
-Changing to the ``gauged frame'' ({\bf c},{\bf d}) makes the drive diagonal in momentum, but comes at the cost of making kinetic energy spin-dependent.
+Changing to the ``gauged frame'' ({\bf c},{\bf d}), essentially by shifting the momentum label $q$ for each spin state $\mu$, makes the drive diagonal in the momentum index, but comes at the cost of making kinetic energy spin-dependent.
 }
@@
 Of course, spin-orbit coupling cannot be ``gauged away'' entirely.
-Making a gauge transformation to simplify the drive comes at the cost of making kinetic energy spin dependent, taking
+Making a gauge transformation to simplify the drive comes at the cost of making kinetic energy spin-dependent, taking
 \begin{align}
@@
 However, $H_{\t{int}}^{(\phi)}$ cannot be exactly reduced to a spin Hamiltonian that addresses the same spin degrees of freedom as $H_{\t{kin}}^{(\phi)}$.
-Nonetheless, for sufficiently weak SOC (small $\phi$) the effect of $H_{\t{int}}^{(\phi)}$ may be approximated by $H_{\t{int}}^{(0)}=H_{\t{int}}$, which has previously been benchmarked for SU(2)-symmetric interactions \cite{he2019engineering, smale2019observation}.
-To ensure that $H_{\t{kin}}^{(\phi)}$ does not become trivial as $\phi\to0$, we can simultaneously take $J\to\infty$ while keeping $J\phi$ constant.
+Nonetheless, for sufficiently weak SOC (small $s\phi$) the effect of $H_{\t{int}}^{(\phi)}$ may be approximated by $H_{\t{int}}^{(0)}=H_{\t{int}}$, which has previously been benchmarked for SU(2)-symmetric interactions \cite{he2019engineering, smale2019observation}.
+To ensure that $H_{\t{kin}}^{(\phi)}$ does not become trivial as $\phi\to0$, we can simultaneously take $J\to\infty$ ($J$ large compared to $U$) while keeping $J\phi$ constant.
 Altogether, the interacting spin Hamiltonian in the gauged frame becomes
@@
 
-We now take a closer look at external control fields, which determine the observables we can access and initial states that we can prepare.
+We now take a closer look at external control fields, which determine the observables we can access and the initial states we can prepare.
 To this end, we consider addressing the atoms with several lasers detuned by $\Delta$ below an electronic ${^1}\t{S}_0\to {^3}\t{P}_0$ transition of individual alkaline-earth-like atoms on a one-dimensional lattice with tight transverse confinement.
@@
 In order to make the drive Hamiltonian in Eq.~\eqref{eq:drive_full} more tractable, we orient the lattice along the $z$ axis, $\v\ell=(0,0,1)$, and consider a three-laser drive with a geometry sketched in Figure \ref{fig:3LD}.
-In this setup, two counter-propagating right-circularly polarized lasers point in directions at an angle $\theta$ with the lattice axis, $\v v_\pm=\pm(0,-\sin\theta,\cos\theta)$, and a third linearly-polarized laser points along $\v v_0=(1,0,0)$.
+In this setup, two counter-propagating right-circularly polarized lasers point along $\v v_\pm=\pm(0,-\sin\theta,\cos\theta)$, at an angle $\theta$ to the lattice axis, and a third linearly-polarized laser points along $\v v_0=(1,0,0)$.
 Choosing $\v v_+$ as the spin quantization axis, we can define new drive amplitudes $\Omega_0\equiv-(\Omega_{\v v_0,+}+\Omega_{\v v_0,-})$ and $\Omega_\pm\equiv\pm\Omega_{\v v_\pm,+}$, as well as a SOC angle $\phi\equiv\kappa\v v_+\c\v\ell=\kappa\cos\theta$, in terms of which the drive Hamiltonian becomes
@@
 \caption{
-Drive Hamiltonians (left column) that can be implemented with different amplitude-matching conditions (right column), some of which are specified by a sign $\sigma\in\set{+1,-1}$.
-The drives shown here are equal to that Eq.~\eqref{eq:drive_single} up to a possible energy shift of $s_\x^2+s_\y^2+s_\z^2=s(s+1)$, and come in mutually commuting pairs: a drive with $\Omega_m=1$ and $\Omega_n=0$ for both $n\ne m$ commutes with the drive in which both $\abs*{\Omega_n}=1$ but $\Omega_m=0$.
+Drive Hamiltonians (left column) that can be implemented with different amplitude-matching conditions (right column), some of which are specified by an arbitrary sign $\sigma\in\set{+1,-1}$.
+The drives shown here are equal to that Eq.~\eqref{eq:drive_single} up to a possible energy shift of $s_\x^2+s_\y^2+s_\z^2=s(s+1)$, and come in mutually commuting pairs: a drive with $\Omega_m=1$ and $\Omega_n=0$ for both $n\ne m$ commutes with the drive in which $\Omega_m=0$ and both $\abs*{\Omega_n}=1$.
 }
@@
 Initial states are nominally prepared in the ``lab frame'', and must be transformed according to the gauge transformation $c_{q\mu}^\dag\to c_{q-\mu\phi,\mu}^\dag$ prior to evolution under the ``gauged frame'' spin Hamiltonian $H_{\t{spin}}$ in Eq.~\eqref{eq:H_spin} or the three-laser drive $H_{\t{3LD}}$ in Eq.~\eqref{eq:drive_all}.
-We assume the capability to prepare an initial state in which all spins are maximally polarized along $\v v$, i.e.~$\ket{\v v}^{\otimes N}=\ket{s}^{\otimes N}$, which is unaffected by the gauge transformation (up to a global phase).
+We assume the capability to prepare an initial state in which all spins are maximally polarized along $\v v_+$, i.e.~$\ket{\v v_+}^{\otimes N}=\ket{s}^{\otimes N}$, which is unaffected by the gauge transformation (up to a global phase).
 The three-laser then allows us to rotate this state into one that is polarized along any spatial axis (in the gauged frame).
 In addition, when $n>2$ the three-laser drive allows us to prepare product states with nontrivial intra-spin correlations.
-For example, by choosing $H_{\t{drive}}^{\t{single}}\propto s_\x^2$ (see Table \ref{tab:drives}), we can transform the polarized state $\ket{s}^{\otimes N}$ into $\p{\ket{s}+\ket{-s}}^{\otimes N}$, where every spin is in a coherent superposition of states pointing in opposite directions along $\v v$.
-When $n>2$, this state has a vanishing mean spin vector, $\bk{S_\x}=\bk{S_\y}=\bk{S_\z}=0$ (where $S_\alpha = \sum_j s_{\alpha j}$), but $\bk{S_\z^2}=Ns^2$.
+For example, by choosing $H_{\t{drive}}^{\t{single}}\propto s_\x^2$ (see Table \ref{tab:drives}), we can transform the polarized state $\ket{s}^{\otimes N}$ into $\p{\ket{s}+\ket{-s}}^{\otimes N}$, where every spin is in a coherent superposition of states pointing in opposite directions along $\v v_+$.
+This state has a mean spin vector that rapidly vanishes with a growing spin dimension $n=2s+1$, $\bk{S_\y}=\bk{S_\z}=0$ and $\bk{S_\x}=N\times 2s/2^{2s}=O(s/4^s)$ (where $S_\alpha = \sum_j s_{\alpha j}$), but a growing axial spin variance $\bk{S_\z^2}=Ns^2$.
 
@@
 
-We now return to the spin Hamiltonian $H_{\t{spin}}$ in Eq.~\eqref{eq:H_spin}, which is valid in the weak-SOC limit $s\phi\ll1$, and henceforth work exclusively in the ``gauged frame'' of this $H_{\t{spin}}$ and the three-laser drive $H_{\t{3LD}}$ in Eq.~\eqref{eq:drive_all}.
-At the mean-field (MF) level, the spin Hamiltonian becomes
+We now return to the spin Hamiltonian $H_{\t{spin}}$ in Eq.~\eqref{eq:H_spin}, which is valid in the weak-SOC limit $s\phi\ll1$, and henceforth work exclusively in the ``gauged frame'' of $H_{\t{spin}}$ and the three-laser drive $H_{\t{3LD}}$ in Eq.~\eqref{eq:drive_all}.
+At the mean-field (MF) level, up to constant energy shifts the undriven spin Hamiltonian becomes
 \begin{align}
@@
 \begin{align}
-  \bbk{\O}_\MF = \lim_{T\to\infty} \f1T \int_0^T \dd t\, \bk{\O\p{t}}_\MF,
+  \bbk{\O}_\MF = \lim_{T\to\infty} \f1T \int_0^T \dd t \bk{\O\p{t}}_\MF,
 \end{align}
@@
 \caption{
-Long-time averages of $\bar\sigma_\x$ and $\ss$ for different spin dimensions $n$ (indicated in the legend) as determined by mean-field simulations of $N=100$ spins initially in the $x$-polarized state $\ket\x^{\otimes N}$.
-Insets show same data with a rescaling $J\phi/U\to J\phi/U \times (n/2)^{1/3}$, $\bbk{\bar\sigma_x}_\MF\to\bbk{\bar\sigma_x}_\MF/\gamma\p{n/2}$, and $\bbk{\ss}_\MF\to\bbk{\ss}_\MF/\gamma\p{n}-1$, where $\gamma\p{k}$ is defined in Eq.~\eqref{eq:gamma}.
+Long-time averages of $\bar\sigma_\x$ and $\ss$ for different spin dimensions $n$ (indicated in the legend) as determined by mean-field simulations of $N=100$ spins initially in the $x$-polarized state $\ket\x^{\otimes N}$ up to a time $T=10^5/U$.
+Insets show same data after normalizing $J\phi/U\to J\phi/U \times (n/2)^{1/3}$, as well as transforming $\bbk{\bar\sigma_\x}_\MF$ and $\bbk{\ss}_\MF$ according to Eq.~\eqref{eq:rescale}.
 }
@@
 As expected, the spin model exhibits a mean-field dynamical phase transition between an ordered phase at small $J\phi/U$ and a disordered phase at large $J\phi/U$.
-The ordered phase has a non-zero transverse magnetization $\bbk{\sigma_\x}_\MF$ and a large interaction energy $\bbk{\ss}_\MF$, while the disordered phase has no transverse magnetization and a minimal interaction energy.
+The ordered phase has a non-zero transverse magnetization $\bbk{\sigma_\x}_\MF$ and a large interaction energy $\bbk{\ss}_\MF$, while the disordered phase has no transverse magnetization and an interaction energy close to its minimal value.
 Numerically, we find that the transition between these phases occurs at a critical field $\p{J\phi/U}_{\t{crit}}=\p{n/2}^{-\alpha}$ with $\alpha\approx1/3$ (see Figure \ref{fig:crit_fields_X}).
@@
 
-As shown in insets of Figure \ref{fig:mean_op_ss_X}, mean-field results for different spin dimensions $n$ collapse onto each other when normalizing the field $J\phi/U$ to its critical value, $\p{J\phi/U}\to\p{J\phi/U}\times\p{n/2}^{1/3}$, and rescaling $\bbk{\bar\sigma_x}_\MF\to\bbk{\bar\sigma_x}_\MF/\gamma\p{n/2}$ and $\bbk{\ss}_\MF\to\bbk{\ss}_\MF/\gamma\p{n}-1$, where
+As shown in insets of Figure \ref{fig:mean_op_ss_X}, mean-field results for different spin dimensions $n$ collapse onto each other when normalizing the field $J\phi/U$ to its critical value, $\p{J\phi/U}\to\p{J\phi/U}\times\p{n/2}^{1/3}$, and rescaling
 \begin{align}
-  \gamma\p{k} \equiv \f{\Gamma\p{k-\frac12}}{\sqrt\pi\,\Gamma\p{k}}.
+  \bbk{\bar\sigma_x}_\MF \to \f{\bbk{\bar\sigma_x}_\MF}{\gamma\p{n/2}},
+  &&
+  \bbk{\ss}_\MF \to \f{\bbk{\ss}_\MF-\gamma\p{n}}{1-\gamma\p{n}},
+  \label{eq:rescale}
+\end{align}
+where
+\begin{align}
+  \gamma\p{k} \equiv \f{\Gamma\p{k-\frac12}}{\sqrt\pi\,\Gamma\p{k}}
+  \stackrel{k\ge2}{\approx} \f1{\sqrt{\pi(k-1)}}.
   \label{eq:gamma}
@@
   = \Tr\sp{\p{\diag\op\x}^2}
-  = \gamma\p{n} \approx \f1{\sqrt{2\pi s}}.
+  = \gamma\p{n}.
 \end{align}
-The same result can be obtained by considering the interaction energy of two counter-rotating spins, whose long-time average is essentially the time-averaged overlap of their phase-space distributions (see Figure \ref{fig:spin_dist}).
+The same result can be obtained by computing the long-time-averaged interaction energy of two spins precessing at different rates.
 
 In the weak-field limit $J\phi/U\to0^+$, the spin-locking $\v S\c\v S$ interactions of the spin Hamiltonian $H_{\t{spin}}$ effectively restrict dynamics to the permutationally-symmetric (PS) manifold.
-At second order in perturbation theory, the spin Hamiltonian thus becomes
+At second order in perturbation theory, the effective spin Hamiltonian within the PS manifold becomes (see Appendix \ref{sec:pert_theory})
 \begin{align}
@@
   = \f{\p{J\phi}^2}{\p{N-1}U}
-  \times \sp{S_\z^2 - N\sum_q s_{\z,q}^2}.
+  \times \sp{S_\z^2 - N\sum_q s_{\z,q}^2},
 \end{align}
-However, mean-field theory fails to capture inter-spin correlations that accumulate throughout evolution under $H_{\t{spin}}^{\t{eff}}$.
-For example, $\ss$ is constant (equal to 1) in the PS manifold, but $\lim_{J\phi/U\to0^+}\bbk{\ss}_\MF<1$ when $n>2$ in Figure \ref{fig:mean_op_ss_X}.
-Treating $H_{\t{spin}}^{\t{eff}}$ at the mean-field level would also yield a homogeneous Hamiltonian that conserves $\ss$, which is again inconsistent with the weak-field limits of $\bbk{\ss}_\MF$.
-This inconsistency between mean-field theories stems from the fact that applying perturbation theory does not commute with a mean-field factorization of spin operators.
-In order to reproduce the weak-field asymptotic values of $\bbk{\bar\sigma_\x}_\MF$ and $\bbk{\ss}_\MF$, therefore, we must instead consider the mean-field Hamiltonian $H_\MF$ in Eq.~\eqref{eq:H_MF}.
+whose mean-field version is
+\begin{align}
+  H_{\t{MFT}}^{\t{eff}} = -\f{\p{J\phi}^2}{U} \sum_q s_{\z,q}^2,
+\end{align}
+where we have used the fact that the axial magnetizations $\bk{s_{\z,q}}=0$ are conserved.
+The weak-field effective Hamiltonian thus preserves permutational symmetry, so $\bbk{\ss}_\MF\to1$ as $J\phi/U\to0^+$, while the transverse magnetization $\bbk{\bar\sigma_\x}_\MF$ approaches the long-time average of $\sigma_\x\equiv s_\x/s$ for a single spin:
+\begin{align}
+  \lim_{J\phi/U\to0^+} \bbk{\bar\sigma_\z}_\MF
+  = \lim_{T\to\infty} \f1T \int_0^T \dd\tau
+  \bk{\x|\sigma_\x\p{\tau}|\x},
+\end{align}
+where
+\begin{align}
+  \sigma_\x\p{\tau} = e^{\i\tau s_\z^2} \sigma_\x e^{-\i\tau s_\z^2}.
+\end{align}
+By mapping the Hilbert space of an $n$-level spin to the Dicke (PS) manifold of $n-1$ spin-$\frac12$ particles, we can adapt exact analytical results for the Ising model \cite{foss-feig2013nonequilibrium}\footnote{See Appendix K of Ref.~\cite{perlin2020shorttime} for a simpler adaptation of the analytics in Ref.~\cite{foss-feig2013nonequilibrium} to the one-axis twisting model $H_{\t{OAT}}=\chi s_\z^2$.} to find that
+\begin{align}
+  \bk{\x|\sigma_\x\p{\tau}|\x} = \p{\cos\tau}^{n-2},
+\end{align}
+so for even $n$
+\begin{align}
+  \lim_{J\phi/U\to0^+} \bbk{\bar\sigma_\z}_\MF
+  = \f1{2\pi} \int_0^{2\pi} \dd\tau \p{\cos\tau}^{n-2}
+  = \gamma\p{\f{n}{2}}.
+\end{align}
 
@@
 \appendix
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Perturbation theory for SU($n$) ferromagnets}
+\label{sec:pert_theory}
+
+\red{[to be completed]}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Restricting spin operators to the permutationally-symmetric manifold}
+\label{sec:PS_ops}
+
+Here we provide the restriction of a general $M$-body spin operator $\O$ to the permutationally-symmetric (PS) manifold of $N$ spins (each with $n$ internal states).
+Denoting the projector onto the PS manifold by $\P$, our task is essentially to find the coefficients of the expansion
+\begin{align}
+  \P \O_M \P = \sum_{a,b\in\A_n\p{N}} \bk{a|\O_M|b} \op{a}{b},
+\end{align}
+where $\A_n\p{N}$ is the set of all ways to assign $N$ (identical) spins to $n$ (distinct) states, such that for any $a\in\A_n\p{N}$ the state $\ket{a}=\ket{a_1,a_2,\cdots,a_n}$ is labeled by the occupation number $a_\mu$ of state $\mu$, with $\sum_\mu a_\mu=N$.
+Written out explicitly,
+\begin{align}
+  \ket{a} = \f1{\sqrt{\C\p{a}}}
+  \sum_{\substack{\t{distinct}\\\t{permutations}\\\Pi\,\t{of}\,\tilde a}}
+  \Pi\ket{\tilde a},
+  &&
+  \ket{\tilde a} \equiv \bigotimes_\mu \ket{\mu}^{\otimes a_\mu},
+  &&
+  \C\p{a} \equiv \f{\p{\sum_\mu a_\mu}!}{\prod_\nu a_\nu!}.
+\end{align}
+Here $\C\p{a}$ is a multinomial coefficient that counts the number of distinct ways to permute the tensor factors of the ``standard-ordered'' state $\ket{\tilde a}$, enforcing $\bk{a|a}=1$.
+Using these states, with some combinatorics we can expand
+\begin{align}
+  \bk{a|\O_M|b} =
+  \sum_{\substack{\alpha,\beta\in\A_n\p{M}\\\alpha\le a,\beta\le b}}
+  \delta_{a-\alpha,b-\beta}
+  \sqrt{\f{\C\p{\alpha}\C\p{a-\alpha}\C\p{\beta}\C\p{b-\beta}}
+    {\C\p{a}\C\p{b}}}
+  \, \bk{\alpha|\O_M|\beta},
+  \label{eq:multi_body_eval}
+\end{align}
+where the restriction $\alpha\le a$ and the difference $a-\alpha$ are evaluated element-wise, i.e.~$\alpha\le a\implies\alpha_\mu\le a_\mu$ and $\p{a-\alpha}_\mu=a_\mu-\alpha_\mu$ for all $\mu$; and $\delta_{cd}=1$ if $c=d$ and zero otherwise.
+We sum over both $\alpha$ and $\beta$ above merely to keep the expression symmetric with respect to transposition $\p{a,\alpha}\leftrightarrow\p{b,\beta}$; in practice, one can simply sum over $\alpha\in\A_n\p{M}$ and set $\beta=b-a+\alpha$, throwing out terms with any $\beta_\mu<0$.
+Note that, by slight abuse of notation, the operator $\O_M$ on the left of \eqref{eq:multi_body_eval} acts on an arbitrary choice of $M$ spins (out of $N$), whereas the operator $\O_M$ on the right of \eqref{eq:multi_body_eval} is simply an $M$-spin operator, with matrix elements $\bk{\alpha|\O_M|\beta}$ evaluated with respect to the PS $M$-spin states $\ket\alpha,\ket\beta\in\A_n\p{M}$.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Schwinger boson equations of motion}
+\section{Schwinger boson equations of motion for quadratic spin Hamiltonians}
 \label{sec:bosons}

### 2021-04-01 collective_sun/sun_perturbation.tex
fix minor error in SU(n) perturbation theory notes

@@
 where the restriction $\alpha\le a$ and difference $a-\alpha$ are
-evaluated element-wise, i.e.~$\alpha<a\implies \alpha_\mu\le a_\mu$
+evaluated element-wise, i.e.~$\alpha\le a\implies \alpha_\mu\le a_\mu$
 and $\p{a-\alpha}_\mu=a_\mu-\alpha_\mu$ for all $\mu$; and

### 2021-04-01 sun_phases/spin_model_paper/main.tex
add back the *correct* 'back-to-back' / 'XX' state

@@
 
-If the interaction energy $U_0$ is smaller than the single-particle bandwidth $\sim J$, then the mode-changing collisions in $H_{\t{int}}$ become off-resonant, motivating the frozen-mode approximation $\set{k,p}=\set{\ell,q}$ (which we assume for the remainder of this work).
+If the interaction energy $U_0$ is smaller than the single-particle bandwidth $\sim J$, then the mode-changing collisions in $H_{\t{int}}$ become off-resonant, motivating the frozen-mode approximation $\set{k,p}=\set{\ell,q}$ (which we assume for the remainder of this work)\footnote{Note that the frozen-mode approximation neglects correlated momentum-hopping terms of the form $c_{\pi-p,\mu}^\dag c_{\pi-q,\mu} c_{p\nu}^\dag c_{q\nu}$, which conserve both momentum and energy.
+We defer a careful treatment of these terms to future work, noting only that they vanish on the manifold of permutationally-symmetric spin states with one atom per lattice site.}.
 The operator content of terms with $k=\ell$ and $p=q$ takes the form $c_{\ell\mu}^\dag c_{\ell\mu} c_{q\nu}^\dag c_{q\nu}$, so these terms contribute an overall energy $\frac12NU$ that can be neglected in the absence of coherent atom number fluctuations.
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{spin_model.pdf}
+\caption{
+(a) Ultracold atoms on a lattice of $L$ sites can tunnel between neighboring lattice sites at a rate $J$, and locally repel each other with interaction energy $U_0$.
+(b,c) When the interaction energy $U_0$ is smaller than the single-particle bandwidth $\sim J$, the frozen-mode approximation enables writing the interaction Hamiltonian as a spin model consisting of exchange terms $\v s_p\c\v s_q$, which swap the states two spins pinned to momentum modes $p,q$.
+(d) Interactions thereby open an energy gap $U=U_0\times N/L$ between the manifold of permutationally-symmetric states of $N$ spins, and the orthogonal complement of states that break spin-permutation symmetry.
+}
+\label{fig:spin_model}
+\end{figure}
+
 We now discuss the spin Hamiltonian $H_{\t{int}}$ in Eq.~\eqref{eq:H_int}.
@@
 In this sense, $\v s_p\c\v s_q$ is analogous to the enforcement of SU(2) spin alignment by ferromagnetic interactions, which similarly assigns (distinct) definite energies to anti-symmetric spin-0 singlet $\ket{\up\dn}-\ket{\dn\up}$ and the symmetric spin-1 triplets $\set{\ket{\up\up},\ket{\dn\dn},\ket{\up\dn}+\ket{\dn\up}}$.
+By summing over all pair-wise exchange terms $\v s_p\c\v s_q$, the interaction Hamiltonian $H_{\t{int}}$ energetically enforces a permutational symmetry among all spins, opening an energy gap $U$ between the manifold of all permutationally-symmetric (PS) states and the orthogonal complement of excited (e.g.~spin-wave) states that break permutational symmetry.
+See Figure \ref{fig:spin_model} for a summary of this section thus far.
 
-By summing over all pair-wise exchange terms $\v s_p\c\v s_q$, the interaction Hamiltonian $H_{\t{int}}$ energetically enforces a permutational symmetry among all spins, opening an energy gap $U$ between the manifold of all permutationally-symmetric (PS) states and the orthogonal complement of excited (e.g.~spin-wave) states that break permutational symmetry.
 In the case of SU(2), the PS manifold is precisely the Dicke manifold of collective states $\ket{m_\z}$ with total spin $S=\frac{N}{2}$ and definite spin projection $m_\z\in\set{S,S-1,\cdots,-S}$ onto a fixed quantization axis.
@@
 Simulating dynamics within the PS manifold requires calculating matrix elements $\bk{\ell|\O|m}$ of spin operators $\O$ with respect to PS states $\ket\ell,\ket{m}$; we discuss this calculation in Appendix \ref{sec:PS_ops}.
-
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{spin_model.pdf}
-\caption{
-(a) Ultracold atoms on a lattice can tunnel between neighboring lattice sites at a rate $J$, and locally repel each other with interaction energy $U_0$.
-(b,c) When the interaction energy $U_0$ is smaller than the single-particle bandwidth $\sim J$, the frozen-mode approximation enables writing the interaction Hamiltonian as a spin model consisting of exchange terms $\v s_p\c\v s_q$, which swap the states two spins pinned to momentum modes $p,q$.
-(d) Interactions thereby open an energy gap $U$ between the manifold of permutationally-symmetric states, and the orthogonal complement of states that break permutational symmetry.
-}
-\label{fig:spin_model}
-\end{figure}
 
@@
 
-In principle, making the gauge transformation $c_{q\mu}^\dag\to c_{q-\mu\phi,\mu}^\dag$ also transforms the interaction Hamiltonian as
+In principle, making the gauge transformation $c_{q\mu}^\dag\to c_{q-\mu\phi,\mu}^\dag$ also transforms the interaction Hamiltonian as $H_{\t{int}} \to H_{\t{int}}^{(\phi)}$, where
 \begin{align}
-  H_{\t{int}} \to H_{\t{int}}^{(\phi)}
+  H_{\t{int}}^{(\phi)}
   \equiv -\f{U}{2N} \sum_{p,q,\mu,\nu}
@@
 In addition, when $n>2$ the three-laser drive allows us to prepare product states with nontrivial intra-spin correlations.
-For example, by choosing $H_{\t{drive}}^{\t{single}}\propto s_\x^2$ (see Table \ref{tab:drives}), we can transform the polarized state $\ket{s}^{\otimes N}$ into $\p{\ket{s}+\ket{-s}}^{\otimes N}$, where every spin is in a coherent superposition of states pointing in opposite directions along $\v v_+$.
-This state has a mean spin vector that rapidly vanishes with a growing spin dimension $n=2s+1$, $\bk{S_\y}=\bk{S_\z}=0$ and $\bk{S_\x}=N\times 2s/2^{2s}=O(s/4^s)$ (where $S_\alpha = \sum_j s_{\alpha j}$), but a growing axial spin variance $\bk{S_\z^2}=Ns^2$.
+For example, we can prepare an $N$-fold product of the state
+\begin{align}
+  \ket{\x\x} \equiv e^{-\i\frac{\pi}{2}s_\y^2} \ket{s}
+  \simeq \ket{+\x} + \ket{-\x},
+\end{align}
+where $\simeq$ denotes equality up to a global phase and normalization factor, and
+\begin{align}
+  \ket{\pm\x} \equiv e^{\mp\i\frac{\pi}{2}s_\y} \ket{s}
+  = \f1{2^s} \sum_\mu \p{\pm1}^{s-\mu}
+  { 2s \choose s+\mu }^{1/2} \ket{\mu}
+\end{align}
+are states of a single spin polarized in opposite directions along the $x$ axis.
+Here ${m \choose k}$ is a binomial coefficient.
+The state $\ket{\x\x}$ has a vanishing mean spin vector, $\bk{S_\x}=\bk{S_\y}=\bk{S_\z}=0$, (where $S_\alpha = \sum_j s_{\alpha j}$), but variances $\bk{S_\x^2}=N\times s^2$ and $\bk{S_\y^2}=\bk{S_\z^2}=N\times s/2$.
 
@@
 \caption{
-Long-time averages of $\bar\sigma_\x$ and $\ss$ for different spin dimensions $n$ (indicated in the legend) as determined by mean-field simulations of $N=100$ spins initially in the $x$-polarized state $\ket\x^{\otimes N}$ up to a time $T=10^5/U$.
+Long-time averages of $\bar\sigma_\x$ and $\ss$ for different spin dimensions $n$ (indicated in the legend) as determined by mean-field simulations of $N=100$ spins initially in the $x$-polarized state $\ket{+\x}^{\otimes N}$ up to a time $T=10^5/U$.
 Insets show same data after normalizing $J\phi/U\to J\phi/U \times (n/2)^{1/3}$, as well as transforming $\bbk{\bar\sigma_\x}_\MF$ and $\bbk{\ss}_\MF$ according to Eq.~\eqref{eq:rescale}.
@@
 
-Figure \ref{fig:mean_op_ss_X} shows the long-time averages of the transverse magnetization $\bar\sigma_\x$ and interaction energy $\ss$ as computed by mean-field simulations of $N=100$ spins in an initial $x$-polarized state $\ket\x^{\otimes N}$, where
-\begin{align}
-  \ket\x \equiv e^{-\i\frac{\pi}{2}s_\y}\ket{s}
-  = \f1{2^s} \sum_\mu { 2s \choose s+\mu }^{1/2} \ket{\mu}.
-  \label{eq:state_x}
-\end{align}
-Here ${m \choose k}$ is a binomial coefficient.
+Figure \ref{fig:mean_op_ss_X} shows the long-time averages of the transverse magnetization $\bar\sigma_\x$ and interaction energy $\ss$ as computed by mean-field simulations of $N=100$ spins in an initial $x$-polarized state $\ket{+\x}^{\otimes N}$.
 As expected, the spin model exhibits a mean-field dynamical phase transition between an ordered phase at small $J\phi/U$ and a disordered phase at large $J\phi/U$.

### 2021-04-01 sun_phases/scripts/compute_boson_states.py

@@ -3,7 +3,7 @@
 import os, sys, time, scipy
 import numpy as np
 
-from dicke_methods import spin_op_z_dicke, spin_op_y_dicke
+from dicke_methods import spin_op_z_dicke, spin_op_y_dicke, coherent_spin_state
 from boson_methods import polarized_state, boson_mft_state, evolve_mft, \
     compute_avg_var_vals
 
@@ -12,7 +12,7 @@ spin_dim = int(sys.argv[2])
 spin_num = int(sys.argv[3])
 log10_field = sys.argv[4]
 
-assert( init_state_str in [ "X" ] )
+assert( init_state_str in [ "X", "XX" ] )
 assert( spin_dim % 2 == 0 )
 assert( spin_num % 2 == 0 )
 
@@ -41,7 +41,13 @@ def spin_coeff(qq):
 field_data = [ sz, field * spin_coeff(np.arange(spin_num)) ]
 
 # construct initial state
-init_state = polarized_state("+X", spin_dim, spin_num)
+if init_state_str == "X":
+    init_state = polarized_state("+X", spin_dim, spin_num)
+elif init_state_str == "XX":
+    sy = spin_op_y_dicke(spin_dim-1)
+    state_z = coherent_spin_state("+Z", spin_dim-1)
+    state_xx = scipy.sparse.linalg.expm_multiply(-1j * np.pi/2 * sy @ sy, state_z)
+    init_state = boson_mft_state(state_xx, spin_dim, spin_num)
 
 prev_state = init_state # state at last simulated time
 prev_mean = 0 # running time-averaged state

### 2021-04-01 sun_phases/scripts/plot_boson_states.py

@@ -14,7 +14,7 @@ from boson_methods import extract_avg_var_state
 init_state_str = sys.argv[1]
 spin_num = int(sys.argv[2])
 
-assert( init_state_str in [ "X" ] )
+assert( init_state_str in [ "X", "XX" ] )
 
 data_dir = "../data/spin_bosons/"
 fig_dir = "../spin_model_paper/figures/"
@@ -50,7 +50,10 @@ def gamma(kk):
 def spin_op(dim):
     spin = (dim-1)/2
     op = spin_op_x_dicke(dim-1).todense() / spin
+    if init_state_str == "XX": op = op @ op
     return np.array(op)
+op_str = r"\bar\sigma_{\mathrm{x}}"
+if init_state_str == "XX": op_str += r"^2"
 
 inset = mpl_toolkits.axes_grid1.inset_locator.inset_axes
 figure, axes = plt.subplots(2, figsize = (2.6,3), sharex = True, sharey = True)
@@ -112,7 +115,7 @@ for dim_idx, dim in enumerate(dims):
     sub_axes[1].semilogx(fields_reduced, mean_ss_inset, ".", **kwargs)
 
 # label axes and set axis ticks
-axes[0].set_ylabel(r"$\bbk{\bar\sigma_{\mathrm{x}}}_\MF$")
+axes[0].set_ylabel(r"$\bbk{" + op_str + r"}_\MF$")
 axes[1].set_ylabel(r"$\bbk{\bar{\bm s}\cdot\bar{\bm s}}_\MF$")
 axes[1].set_xlabel(r"$J\phi/U$")
 axes[1].set_xlim(0.1,10)

### 2021-04-02 sun_phases/spin_model_paper/main.tex
update SU(n) spin model paper to reflect changes to figures

@@
 Here ${m \choose k}$ is a binomial coefficient.
-The state $\ket{\x\x}$ has a vanishing mean spin vector, $\bk{S_\x}=\bk{S_\y}=\bk{S_\z}=0$, (where $S_\alpha = \sum_j s_{\alpha j}$), but variances $\bk{S_\x^2}=N\times s^2$ and $\bk{S_\y^2}=\bk{S_\z^2}=N\times s/2$.
+The state $\ket{\x\x}$ has a vanishing mean spin vector, $\bk{S_\x}=\bk{S_\y}=\bk{S_\z}=0$ (where $S_\alpha = \sum_j s_{\alpha j}$), but variances $\bk{S_\x^2}=N\times s^2$ and $\bk{S_\y^2}=\bk{S_\z^2}=N\times s/2$.
 
@@
 \centering
-\includegraphics{mean_op_ss_X.pdf}
+\includegraphics{mean_sx_ss.pdf}
 \caption{
@@
 }
-\label{fig:mean_op_ss_X}
+\label{fig:mean_sx_ss}
 \end{figure}
 
-Figure \ref{fig:mean_op_ss_X} shows the long-time averages of the transverse magnetization $\bar\sigma_\x$ and interaction energy $\ss$ as computed by mean-field simulations of $N=100$ spins in an initial $x$-polarized state $\ket{+\x}^{\otimes N}$.
+Figure \ref{fig:mean_sx_ss} shows the long-time averages of the transverse magnetization $\bar\sigma_\x$ and interaction energy $\ss$ as computed by mean-field simulations of $N=100$ spins in an initial $x$-polarized state $\ket{+\x}^{\otimes N}$.
 As expected, the spin model exhibits a mean-field dynamical phase transition between an ordered phase at small $J\phi/U$ and a disordered phase at large $J\phi/U$.
-The ordered phase has a non-zero transverse magnetization $\bbk{\sigma_\x}_\MF$ and a large interaction energy $\bbk{\ss}_\MF$, while the disordered phase has no transverse magnetization and an interaction energy close to its minimal value.
-Numerically, we find that the transition between these phases occurs at a critical field $\p{J\phi/U}_{\t{crit}}=\p{n/2}^{-\alpha}$ with $\alpha\approx1/3$ (see Figure \ref{fig:crit_fields_X}).
+The ordered phase has a non-zero transverse magnetization $\bbk{\sigma_\x}_\MF$ and an interaction energy $\bbk{\ss}_\MF$ that asymptotically approaches its maximal value as $J\phi/U\to0^+$.
+The disordered phase has no transverse magnetization, $\bbk{\bar\sigma_\x}_\MF=0$, but an interaction energy $\bbk{\ss}_\MF$ that nonetheless indicates persistent (albeit small) nontrivial inter-spin correlations when $n>2$.
+These nontrivial inter-spin correlations vanish as $J\phi/U\to\infty$, in which case $\bbk{\ss}_\MF$ approaches the minimal value allowed by conservation laws (clarified below).
+By minimizing the reduced field $J\phi/U$ for which $\bbk{\sigma_\x}_\MF=0$, we numerically find that the transition between ordered and disordered phases occurs at a critical field $\p{J\phi/U}_{\t{crit}}=\p{n/2}^{-\alpha}$ with $\alpha\approx1/3$ (see Figure \ref{fig:crit_fields}).
 
@@
 \centering
-\includegraphics{crit_fields_X.pdf}
+\includegraphics{crit_fields.pdf}
 \caption{
@@
 }
-\label{fig:crit_fields_X}
+\label{fig:crit_fields}
 \end{figure}
 
-As shown in insets of Figure \ref{fig:mean_op_ss_X}, mean-field results for different spin dimensions $n$ collapse onto each other when normalizing the field $J\phi/U$ to its critical value, $\p{J\phi/U}\to\p{J\phi/U}\times\p{n/2}^{1/3}$, and rescaling
+As shown in insets of Figure \ref{fig:mean_sx_ss}, mean-field results for different spin dimensions $n$ collapse onto each other when normalizing the field $J\phi/U$ to its critical value, $\p{J\phi/U}\to\p{J\phi/U}\times\p{n/2}^{1/3}$, and rescaling
 \begin{align}
@@
 \end{enumerate*}
-Altogether, the interaction energy $\bbk{\ss}_\MF$ in the strong-field limit is determined by the time-independent diagonal part $\diag\bk{\bar{\v s}\p{t}} = \diag\op\x$, namely
+Altogether, the interaction energy $\bbk{\ss}_\MF$ in the strong-field limit is determined by the time-independent diagonal part $\diag\bk{\bar{\v s}\p{t}} = \diag\op{+\x}$, namely
 \begin{align}
   \lim_{J\phi/U\to\infty} \bbk{\ss}_\MF
-  = \Tr\sp{\p{\diag\op\x}^2}
+  = \Tr\sp{\p{\diag\op{+\x}}^2}
   = \gamma\p{n}.
@@
 \end{align}
-whose mean-field version is
+which at the mean-field level becomes
 \begin{align}
@@
   = \lim_{T\to\infty} \f1T \int_0^T \dd\tau
-  \bk{\x|\sigma_\x\p{\tau}|\x},
+  \bk{+\x|\sigma_\x\p{\tau}|+\x},
 \end{align}
@@
 \end{align}
-By mapping the Hilbert space of an $n$-level spin to the Dicke (PS) manifold of $n-1$ spin-$\frac12$ particles, we can adapt exact analytical results for the Ising model \cite{foss-feig2013nonequilibrium}\footnote{See Appendix K of Ref.~\cite{perlin2020shorttime} for a simpler adaptation of the analytics in Ref.~\cite{foss-feig2013nonequilibrium} to the one-axis twisting model $H_{\t{OAT}}=\chi s_\z^2$.} to find that
+By mapping the Hilbert space of an $n$-level spin onto the PS (Dicke) manifold of $n-1$ spin-$\frac12$ particles, we can adapt exact analytical results for the Ising model \cite{foss-feig2013nonequilibrium}\footnote{See Appendix K of Ref.~\cite{perlin2020shorttime} for a simpler adaptation of the analytics in Ref.~\cite{foss-feig2013nonequilibrium} to the one-axis twisting model $H_{\t{OAT}}=\chi s_\z^2$.} to find that
 \begin{align}
-  \bk{\x|\sigma_\x\p{\tau}|\x} = \p{\cos\tau}^{n-2},
+  \bk{+\x|\sigma_\x\p{\tau}|+\x} = \p{\cos\tau}^{n-2},
 \end{align}

### 2021-04-04 sun_phases/spin_model_paper/main.tex
removing old unused codes fixing some confusions related to phase ambiguities in the definitions of polarized states

@@
 \end{align}
-Here $\v S$ is a collective spin matrix, analogous to the collective spin vector $\vec S=\p{S_\x,S_\y,S_\z}$ in the case of SU(2) \cite{he2019engineering}, with $\v S\c\v S \simeq 2 \vec S\c\vec S = 2 \sum_{\alpha\in\set{\x,\y,\z}}S_\alpha^2$ when $n=2$ (with $\simeq$ denoting equality after modding out identity terms).
+Here $\v S$ is a collective spin matrix, analogous to the collective spin vector $\vec S=\p{S_\x,S_\y,S_\z}$ in the case of SU(2) \cite{he2019engineering}, with $\v S\c\v S \simeq 2 \vec S\c\vec S = 2 \sum_{\alpha\in\set{\x,\y,\z}}S_\alpha^2$ when $n=2$ (here $\simeq$ denotes equality after modding out identity terms).
 
@@
 In addition, when $n>2$ the three-laser drive allows us to prepare product states with nontrivial intra-spin correlations.
-For example, we can prepare an $N$-fold product of the state
+For example, when $n$ is even we can prepare an $N$-fold product of the state
 \begin{align}
-  \ket{\x\x} \equiv e^{-\i\frac{\pi}{2}s_\y^2} \ket{s}
-  \simeq \ket{+\x} + \ket{-\x},
+  e^{-\i\frac{\pi}{2}\p{s_\y+s_\y^2}} \ket{s} \simeq \ket{s} + \ket{-s},
 \end{align}
-where $\simeq$ denotes equality up to a global phase and normalization factor, and
-\begin{align}
-  \ket{\pm\x} \equiv e^{\mp\i\frac{\pi}{2}s_\y} \ket{s}
-  = \f1{2^s} \sum_\mu \p{\pm1}^{s-\mu}
-  { 2s \choose s+\mu }^{1/2} \ket{\mu}
-\end{align}
-are states of a single spin polarized in opposite directions along the $x$ axis.
-Here ${m \choose k}$ is a binomial coefficient.
-The state $\ket{\x\x}$ has a vanishing mean spin vector, $\bk{S_\x}=\bk{S_\y}=\bk{S_\z}=0$ (where $S_\alpha = \sum_j s_{\alpha j}$), but variances $\bk{S_\x^2}=N\times s^2$ and $\bk{S_\y^2}=\bk{S_\z^2}=N\times s/2$.
+where $\simeq$ denotes equality up to a global phase and normalization factor.
+This state has a vanishing mean spin vector, $\bk{s_\x}=\bk{s_\y}=\bk{s_\z}=0$, but variances $\bk{s_\x^2}=\bk{s_\y^2}=s/2$ and $\bk{s_\z^2}=s^2$.
 
@@
 Fixing the atom number $N$, the spin Hamiltonian has one free parameter: the reduced field $J\phi/U$ that determines the relative strength of the single-particle and interaction terms.
-For an initial $N$-fold tensor product state $\ket\psi^{\otimes N}$ one should expect distinct dynamical behaviors when $J\phi/U\ll1$, in which case strong spin-locking interactions should give rise to a long-range ordered phase, as opposed to $J\phi/U\gg1$, in which case long-range order should be destroyed by the strong inhomogeneous magnetic field.
+One should therefore expect distinct dynamical behaviors when $J\phi/U\ll1$, in which case strong spin-locking interactions should give rise to a long-range ordered phase, as opposed to $J\phi/U\gg1$, in which case long-range order should be destroyed by the strong inhomogeneous magnetic field.
 
@@
 \begin{align}
-  \bar\sigma_\x \equiv \f{S_\x}{\norm{S_\x}},
+  \bar\sigma_\x \equiv \f1{Ns} \sum_q s_{\x,q},
   &&
-  \ss = \f{\v S\c\v S}{N^2},
+  \ss = \f{\v S\c\v S}{N^2}.
 \end{align}
-where $\norm{S_\x}=Ns$ is the operator norm of $S_\x$.
-By design, these observables to have norm $\norm{\bar\sigma_\x}=\norm{\ss}=1$ independent of the system size $N$ or spin dimension $n$.
+By design, these observables to have operator norm $\norm{\bar\sigma_\x}=\norm{\ss}=1$ independent of the system size $N$ or spin dimension $n$.
+In the remainder of this section we will assume that $n$ is even, both for the sake of experimental relevance (most relevant atomic nuclei are fermionic) and to avoid complications from parity effects.
 
@@
 \caption{
-Long-time averages of $\bar\sigma_\x$ and $\ss$ for different spin dimensions $n$ (indicated in the legend) as determined by mean-field simulations of $N=100$ spins initially in the $x$-polarized state $\ket{+\x}^{\otimes N}$ up to a time $T=10^5/U$.
+Long-time averages of $\bar\sigma_\x$ and $\ss$ for different spin dimensions $n$ (indicated in the legend) as determined by mean-field simulations of $N=100$ spins initially in the $x$-polarized state $\ket{\x}^{\otimes N}$ up to a time $T=10^5/U$.
 Insets show same data after normalizing $J\phi/U\to J\phi/U \times (n/2)^{1/3}$, as well as transforming $\bbk{\bar\sigma_\x}_\MF$ and $\bbk{\ss}_\MF$ according to Eq.~\eqref{eq:rescale}.
@@
 
-Figure \ref{fig:mean_sx_ss} shows the long-time averages of the transverse magnetization $\bar\sigma_\x$ and interaction energy $\ss$ as computed by mean-field simulations of $N=100$ spins in an initial $x$-polarized state $\ket{+\x}^{\otimes N}$.
+Figure \ref{fig:mean_sx_ss} shows the long-time averages of the transverse magnetization $\bar\sigma_\x$ and interaction energy $\ss$ as computed by mean-field simulations of $N=100$ spins in an initial $x$-polarized state $\ket{\x}^{\otimes N}$, where
+\begin{align}
+  \ket{\x} \equiv e^{-\i\frac{\pi}{2}s_\y} \ket{s}
+  = \f1{2^s} \sum_\mu { 2s \choose s+\mu }^{1/2} \ket{\mu}.
+\end{align}
+Here ${ m \choose k }$ is a binomial coefficient.
 As expected, the spin model exhibits a mean-field dynamical phase transition between an ordered phase at small $J\phi/U$ and a disordered phase at large $J\phi/U$.
@@
 \end{enumerate*}
-Altogether, the interaction energy $\bbk{\ss}_\MF$ in the strong-field limit is determined by the time-independent diagonal part $\diag\bk{\bar{\v s}\p{t}} = \diag\op{+\x}$, namely
+Altogether, the interaction energy $\bbk{\ss}_\MF$ in the strong-field limit is determined by the time-independent diagonal part $\diag\bk{\bar{\v s}}_\MF = \diag\op{\x}$, namely
 \begin{align}
   \lim_{J\phi/U\to\infty} \bbk{\ss}_\MF
-  = \Tr\sp{\p{\diag\op{+\x}}^2}
+  = \Tr\sp{\p{\diag\op{\x}}^2}
   = \gamma\p{n}.
@@
   = \lim_{T\to\infty} \f1T \int_0^T \dd\tau
-  \bk{+\x|\sigma_\x\p{\tau}|+\x},
+  \bk{\x|\sigma_\x\p{\tau}|\x},
 \end{align}
@@
 \begin{align}
-  \bk{+\x|\sigma_\x\p{\tau}|+\x} = \p{\cos\tau}^{n-2},
+  \bk{\x|\sigma_\x\p{\tau}|\x} = \p{\cos\tau}^{n-2},
 \end{align}
@@
 \begin{align}
-  \lim_{J\phi/U\to0^+} \bbk{\bar\sigma_\z}_\MF
+  \lim_{J\phi/U\to0^+} \bbk{\bar\sigma_\x}_\MF
   = \f1{2\pi} \int_0^{2\pi} \dd\tau \p{\cos\tau}^{n-2}

### 2021-04-04 sun_phases/scripts/compute_spin_pair.py

@@ -1,93 +1,0 @@
-#!/usr/bin/env python3
-
-import os, sys, time
-import numpy as np
-import scipy, scipy.integrate
-
-from dicke_methods import spin_op_vec_dicke
-
-np.set_printoptions(linewidth = 200)
-
-dim = sys.argv[1]
-log10_field = sys.argv[2]
-
-sim_time = 10**4
-time_step = 0.1
-ivp_tolerance = 2.220446049250313e-14 # smallest value allowed
-
-data_dir = f"../data/spin_pair/"
-if not os.path.isdir(data_dir):
-    os.makedirs(data_dir)
-
-sys_tag = f"d{dim}_h{log10_field}"
-def data_file(tag):
-    return data_dir + f"{tag}_{sys_tag}.txt"
-
-genesis = time.time()
-####################
-
-dim = int(dim)
-log10_field = float(log10_field)
-
-spin = (dim-1)/2
-field = 10**log10_field / (2*spin)
-
-# spin matrices
-sz, sx, sy = spin_op_vec_dicke(dim-1)
-
-def rank(state):
-    return int(np.round(np.log(state.size)/np.log(dim)))
-
-# join two states into a 1-D array
-def join(S0, S1):
-    return np.array([ S0, S1 ]).ravel()
-
-# split a 1-D array into two matrices
-def split(state):
-    S0 = state[:state.size//2]
-    S1 = state[state.size//2:]
-    S0.shape = S1.shape = (dim,) * rank(S0)
-    return S0, S1
-
-def val(op, state):
-    if rank(state) == 2:
-        return np.trace(op @ state)
-    else:
-        return state.conj() @ ( op @ state )
-
-# time derivative of spin state
-def time_deriv(state, field):
-    B0, B1 = split(state)
-    deriv_0 = B1 * ( B1.conj() @ B0 ) + field * ( sz @ B0 )
-    deriv_1 = B0 * ( B0.conj() @ B1 ) - field * ( sz @ B1 )
-    return -1j * join(deriv_0, deriv_1)
-
-def evolve(initial_state, time_deriv, times, tol = ivp_tolerance):
-    def ivp_deriv(time, state): return time_deriv(state, field)
-    ivp_args = [ ivp_deriv, (times[0], times[-1]), initial_state ]
-    ivp_kwargs = dict( rtol = tol, atol = tol, t_eval = times )
-    ivp_solution = scipy.integrate.solve_ivp(*ivp_args, **ivp_kwargs)
[…]

### 2021-04-04 sun_phases/scripts/plot_spin_pair.py

@@ -1,95 +1,0 @@
-#!/usr/bin/env python3
-
-import os, glob, qutip
-import numpy as np
-import matplotlib.pyplot as plt
-
-from dicke_methods import spin_op_vec_dicke, plot_dicke_state
-
-data_dir = "../data/spin_pair/"
-fig_dir = "../figures/spin_pair/spin_vecs/"
-
-figsize = (3,3)
-
-# set fonts and use latex packages
-params = { "font.family" : "serif",
-           "font.serif" : "Computer Modern",
-           "text.usetex" : True,
-           "font.size" : 10 }
-plt.rcParams.update(params)
-
-##################################################
-
-def get_info(file, string = False):
-    _, dim, log10_field = os.path.basename(file).split("_")
-    dim = dim[1:]
-    log10_field = log10_field[1:-4]
-    if not string:
-        dim = int(dim)
-        log10_field = float(log10_field)
-    return { "dim" : dim, "log10_field" : log10_field }
-
-files = sorted(glob.glob(data_dir + "*.txt"),
-               key = lambda file : list(get_info(file).values()))
-dims = sorted(set( get_info(file)["dim"] for file in files ))
-
-def spin_op_vec(dim):
-    spin = (dim-1)/2
-    sz, sx, sy = spin_op_vec_dicke(dim-1)
-    return np.array([ sx.todense(), sy.todense(), sz.todense() ]) / spin
-spin_op_vecs = { dim : spin_op_vec(dim) for dim in dims }
-
-##################################################
-
-def fig_name(file, aux_dir, tag):
-    _fig_dir = fig_dir + aux_dir + "/"
-    field_text = file.split("_")[-1][1:-4]
-    if len(field_text) in [ 3, 4 ]:
-        _fig_dir += "sweep/"
-    else:
-        _fig_dir += "zoom/"
-    name = file.replace(data_dir + "states", _fig_dir + f"{tag}")
-    return name.replace(".txt", ".png")
-
-def spin_vec(state):
-    return np.einsum("aij,i,j", spin_op_vecs[state.size], state.conj(), state).real
-
-def make_bloch(spin_vecs, view = None):
-    figure = plt.figure(figsize = figsize)
-    bloch = qutip.Bloch(figure)
-    bloch.add_points(spin_vecs.T)
-    bloch.xlabel = [ "", "" ]
[…]

### 2021-04-05 sun_phases/spin_model_paper/main.tex
change to plotting isotropic magnetization, rather than just that along one axis

@@
 
-To investigate these behaviors quantitatively, we examine long-time-averaged observables of the form
+To investigate these behaviors quantitatively, we examine time-averaged observables of the form
 \begin{align}
@@
 where $\bk{\O\p{t}}_\MF$ is the mean-field value of observable $\O$ at time $t$ (see Appendices \ref{sec:MFT} and \ref{sec:bosons} for details about our mean-field simulations).
-Specifically, we consider long-time averages of the normalized transverse magnetization and interaction energy
+Specifically, we consider the time-averaged magnetization
 \begin{align}
-  \bar\sigma_\x \equiv \f1{Ns} \sum_q s_{\x,q},
+  m_\MF \equiv \abs{\bbk{\vec\sigma}_\MF},
   &&
-  \ss = \f{\v S\c\v S}{N^2}.
+  \vec\sigma \equiv \f1{Ns} \times \p{S_\x,S_\y,S_\z},
 \end{align}
-By design, these observables to have operator norm $\norm{\bar\sigma_\x}=\norm{\ss}=1$ independent of the system size $N$ or spin dimension $n$.
+where $S_\alpha \equiv \sum_q s_{\alpha,q}$, and the time-averaged interaction energy
+\begin{align}
+  \bbk{\ss}_\MF = \f1{N^2} \times \bbk{\v S\c\v S}_\MF.
+\end{align}
+By design, these quantities are non-negative and normalized to a maximal value of one, independent of the system size $N$ or spin dimension $n$.
 In the remainder of this section we will assume that $n$ is even, both for the sake of experimental relevance (most relevant atomic nuclei are fermionic) and to avoid complications from parity effects.
@@
 \centering
-\includegraphics{mean_sx_ss.pdf}
+\includegraphics{mean_mag_int.pdf}
 \caption{
-Long-time averages of $\bar\sigma_\x$ and $\ss$ for different spin dimensions $n$ (indicated in the legend) as determined by mean-field simulations of $N=100$ spins initially in the $x$-polarized state $\ket{\x}^{\otimes N}$ up to a time $T=10^5/U$.
+Time-averaged magnetization $m_\MF$ and interaction energy $\bbk{\ss}_\MF$ for different spin dimensions $n$ (indicated in the legend) as determined by mean-field simulations of $N=100$ spins initially in the $x$-polarized state $\ket\x^{\otimes N}$ up to a time $T=10^5/U$.
 Insets show same data after normalizing $J\phi/U\to J\phi/U \times (n/2)^{1/3}$, as well as transforming $\bbk{\bar\sigma_\x}_\MF$ and $\bbk{\ss}_\MF$ according to Eq.~\eqref{eq:rescale}.
 }
-\label{fig:mean_sx_ss}
+\label{fig:mean_mag_int}
 \end{figure}
 
-Figure \ref{fig:mean_sx_ss} shows the long-time averages of the transverse magnetization $\bar\sigma_\x$ and interaction energy $\ss$ as computed by mean-field simulations of $N=100$ spins in an initial $x$-polarized state $\ket{\x}^{\otimes N}$, where
+Figure \ref{fig:mean_mag_int} shows the time-averages of the magnetization $m_\MF$ and interaction energy $\bbk{\ss}_\MF$ as computed by mean-field simulations of $N=100$ spins in an initial $x$-polarized state $\ket\x^{\otimes N}$, where
 \begin{align}
-  \ket{\x} \equiv e^{-\i\frac{\pi}{2}s_\y} \ket{s}
+  \ket\x \equiv e^{-\i\frac{\pi}{2}s_\y} \ket{s}
   = \f1{2^s} \sum_\mu { 2s \choose s+\mu }^{1/2} \ket{\mu}.
@@
 As expected, the spin model exhibits a mean-field dynamical phase transition between an ordered phase at small $J\phi/U$ and a disordered phase at large $J\phi/U$.
-The ordered phase has a non-zero transverse magnetization $\bbk{\sigma_\x}_\MF$ and an interaction energy $\bbk{\ss}_\MF$ that asymptotically approaches its maximal value as $J\phi/U\to0^+$.
-The disordered phase has no transverse magnetization, $\bbk{\bar\sigma_\x}_\MF=0$, but an interaction energy $\bbk{\ss}_\MF$ that nonetheless indicates persistent (albeit small) nontrivial inter-spin correlations when $n>2$.
+The ordered phase has a non-zero magnetization $m_\MF$ and an interaction energy $\bbk{\ss}_\MF$ that asymptotically approaches its maximal value as $J\phi/U\to0^+$.
+The disordered phase has no (time-averaged) magnetization, $m_\MF=0$, but an interaction energy $\bbk{\ss}_\MF$ that nonetheless indicates persistent nontrivial inter-spin correlations when $n>2$.
 These nontrivial inter-spin correlations vanish as $J\phi/U\to\infty$, in which case $\bbk{\ss}_\MF$ approaches the minimal value allowed by conservation laws (clarified below).
-By minimizing the reduced field $J\phi/U$ for which $\bbk{\sigma_\x}_\MF=0$, we numerically find that the transition between ordered and disordered phases occurs at a critical field $\p{J\phi/U}_{\t{crit}}=\p{n/2}^{-\alpha}$ with $\alpha\approx1/3$ (see Figure \ref{fig:crit_fields}).
+By minimizing the reduced field $J\phi/U$ for which $m_\MF=0$, we numerically find that the transition between ordered and disordered phases occurs at a critical field $\p{J\phi/U}_{\t{crit}}=\p{n/2}^{-\alpha}$ with $\alpha\approx1/3$ (see Figure \ref{fig:crit_fields}).
 
@@
 
-As shown in insets of Figure \ref{fig:mean_sx_ss}, mean-field results for different spin dimensions $n$ collapse onto each other when normalizing the field $J\phi/U$ to its critical value, $\p{J\phi/U}\to\p{J\phi/U}\times\p{n/2}^{1/3}$, and rescaling
+As shown in insets of Figure \ref{fig:mean_mag_int}, mean-field results for different spin dimensions $n$ collapse onto each other when normalizing the field $J\phi/U$ to its critical value, $\p{J\phi/U}\to\p{J\phi/U}\times\p{n/2}^{1/3}$, and rescaling
 \begin{align}
-  \bbk{\bar\sigma_x}_\MF \to \f{\bbk{\bar\sigma_x}_\MF}{\gamma\p{n/2}},
+  m_\MF \to \f{m_\MF}{\gamma\p{n/2}},
   &&
@@
 \end{align}
-The rescaling of transverse magnetization and interaction energy can be understood by considering their limiting behavior as $J\phi/U\to\infty$ or $J\phi/U\to0^+$.
+The rescaling of magnetization and interaction energy can be understood by considering their limiting behavior as $J\phi/U\to\infty$ or $J\phi/U\to0^+$.
 
 In the strong-field limit $J\phi/U\to\infty$, we can ignore interactions and treat spins as though they simply precess at different rates.
-The long-time-averaged transverse magnetization $\bbk{\bar\sigma_\x}_\MF$ then trivially vanishes as $J\phi/U\to\infty$.
+The time-averaged transverse magnetization $m_\MF$ then trivially vanishes as $J\phi/U\to\infty$.
 The interaction energy $\bk{\ss}_\MF = \bk{\bar{\v s}}_\MF \c \bk{\bar{\v s}}_\MF$, meanwhile, has contributions from
@@
 \item the diagonal parts of the mean spin matrix $\bk{\bar{\v s}}_\MF$, which are conserved by inhomogeneous spin precession, and
-\item the off-diagonal parts of $\bk{\bar{\v s}}_\MF$, whose oscillations average to zero when evaluating the long-time average $\bbk{\ss}_\MF$.
+\item the off-diagonal parts of $\bk{\bar{\v s}}_\MF$, whose oscillations average to zero when evaluating the time average in $\bbk{\ss}_\MF$.
 \end{enumerate*}
@@
 \end{align}
-The same result can be obtained by computing the long-time-averaged interaction energy of two spins precessing at different rates.
+The same result can be obtained by computing the time-averaged interaction energy of two spins precessing at different rates.
 
-In the weak-field limit $J\phi/U\to0^+$, the spin-locking $\v S\c\v S$ interactions of the spin Hamiltonian $H_{\t{spin}}$ effectively restrict dynamics to the permutationally-symmetric (PS) manifold.
+In the weak-field limit $J\phi/U\to0^+$, the spin-locking $\v S\c\v S$ interactions of the spin Hamiltonian $H_{\t{spin}}$ energetically restrict dynamics to the permutationally-symmetric (PS) manifold.
 At second order in perturbation theory, the effective spin Hamiltonian within the PS manifold becomes (see Appendix \ref{sec:pert_theory})
@@
 \end{align}
-which at the mean-field level becomes
+whose mean-field version is
 \begin{align}
@@
 where we have used the fact that the axial magnetizations $\bk{s_{\z,q}}=0$ are conserved.
-The weak-field effective Hamiltonian thus preserves permutational symmetry, so $\bbk{\ss}_\MF\to1$ as $J\phi/U\to0^+$, while the transverse magnetization $\bbk{\bar\sigma_\x}_\MF$ approaches the long-time average of $\sigma_\x\equiv s_\x/s$ for a single spin:
+The weak-field effective Hamiltonian preserves permutational symmetry, so $\bbk{\ss}_\MF\to1$ as $J\phi/U\to0^+$.
+In addition, invariance of the initial state $\ket\x^{\otimes N}$ and the effective Hamiltonian $H_{\t{spin}}^{\t{eff}}$ under rotation by $e^{-\i\pi S_\x}$ implies that
 \begin{align}
-  \lim_{J\phi/U\to0^+} \bbk{\bar\sigma_\z}_\MF
-  = \lim_{T\to\infty} \f1T \int_0^T \dd\tau
-  \bk{\x|\sigma_\x\p{\tau}|\x},
+  \bk{S_\y} = \bk{e^{\i\pi S_\x} S_\y e^{-\i\pi S_\x}} = -\bk{S_\y}
+\end{align}
+at all times, so $\bk{S_\y}=0$.
+Together with permutational symmetry, the fact that $\bk{S_\y}=\bk{S_\z}=0$ implies that the magnetization $m_\MF$ is determined by the time-average of $\sigma_\x\equiv s_\x/s$ for a single spin:
+\begin{align}
+  \lim_{J\phi/U\to0^+} m_\MF
+  = \abs{\lim_{T\to\infty} \f1T \int_0^T \dd\tau
+  \bk{\x|\sigma_\x\p{\tau}|\x}},
 \end{align}
@@
 \begin{align}
-  \lim_{J\phi/U\to0^+} \bbk{\bar\sigma_\x}_\MF
-  = \f1{2\pi} \int_0^{2\pi} \dd\tau \p{\cos\tau}^{n-2}
+  \lim_{J\phi/U\to0^+} m_\MF
+  = \abs{\f1{2\pi} \int_0^{2\pi} \dd\tau \p{\cos\tau}^{n-2}}
   = \gamma\p{\f{n}{2}}.

### 2021-04-05 sun_phases/scripts/plot_boson_states.py

@@ -8,7 +8,7 @@ import scipy, scipy.optimize, scipy.special
 import mpl_toolkits
 from mpl_toolkits.axes_grid1.inset_locator import inset_axes
 
-from dicke_methods import spin_op_x_dicke
+from dicke_methods import spin_op_vec_dicke
 from boson_methods import extract_avg_var_state
 
 init_state_str = sys.argv[1]
@@ -65,24 +65,26 @@ for dim_idx, dim in enumerate(dims):
         = zip(*sorted([ ( get_info(file)["log10_field"], file )
                         for file in files if get_info(file)["dim"] == dim ]))
     log10_fields = np.array(log10_fields)
-    log10_fields_reduced = log10_fields + np.log10((dim/2)**(1/3))
     fields = 10**log10_fields
-    fields_reduced = 10**log10_fields_reduced
+    fields_reduced = fields * (dim/2)**(1/3)
 
-    # normalized spin-x operator
+    # normalized spin operator vector
     spin = (dim-1)/2
-    sx_vec = np.array(spin_op_x_dicke(dim-1).todense()).ravel() / spin
+    spin_op_vec = np.array([ np.array(op.todense()).ravel() / spin
+                             for op in spin_op_vec_dicke(dim-1) ]).T
 
     # collect long-time average data
-    mean_sx = np.zeros(log10_fields.size)
-    mean_ss = np.zeros(log10_fields.size)
-    avg_spect = np.zeros((log10_fields.size,dim))
-    var_spect = np.zeros((log10_fields.size,dim**2))
+    mean_mag = np.zeros(fields.size)
+    mean_int = np.zeros(fields.size)
+    avg_spect = np.zeros((fields.size,dim))
+    var_spect = np.zeros((fields.size,dim**2))
+    avg_states = []
     for idx, file in enumerate(dim_files):
         avg_state, var_state \
             = extract_avg_var_state(np.loadtxt(file, dtype = complex), dim)
-        mean_sx[idx] = ( avg_state.ravel().conj() @ sx_vec ).real
-        mean_ss[idx] = np.einsum("mnnm->", var_state).real
+        mean_mag[idx] = np.linalg.norm( avg_state.ravel().conj() @ spin_op_vec )
[…]

### 2021-04-05 sun_phases/spin_model_paper/main.tex
choose 'multi-cat' states to simulate

@@
 \renewcommand{\SS}{\mathbb{S}}
+\newcommand{\ZZ}{\mathbb{Z}}
 \newcommand{\A}{\mathcal{A}}
@@
 \begin{align}
-  H_{\t{spin}} = -\f{U}{2N} \v S\c\v S + J\phi \sum_q \sin\p{q} s_{\z,q}.
+  H_{\t{spin}} = -\f{U}{2N} \v S\c\v S + J\phi \sum_q \sin\p{q} s_{\z,q},
   \label{eq:H_spin}
 \end{align}
+consisting of a spin-locking $\v S\c\v S$ term that energetically favors permutational symmetry, and an inhomogeneous magnetic field that causes inter-spin dephasing.
 
@@
 \begin{align}
-  e^{-\i\frac{\pi}{2}\p{s_\y+s_\y^2}} \ket{s} \simeq \ket{s} + \ket{-s},
+  \frac{\ket{s} + \ket{-s}}{\sqrt{2}}
+  \simeq e^{-\i\frac{\pi}{2}\p{s_\y + s_\y^2}} \ket{s},
 \end{align}
-where $\simeq$ denotes equality up to a global phase and normalization factor.
+where $\simeq$ denotes equality up to a global phase.
 This state has a vanishing mean spin vector, $\bk{s_\x}=\bk{s_\y}=\bk{s_\z}=0$, but variances $\bk{s_\x^2}=\bk{s_\y^2}=s/2$ and $\bk{s_\z^2}=s^2$.
@@
 where $\bar{\v s}\equiv\frac1N\sum_q\v s_q$ is the average spin matrix.
+We assume that all momenta $q\in\ZZ_N\times 2\pi/N$ are occupied.
 Fixing the atom number $N$, the spin Hamiltonian has one free parameter: the reduced field $J\phi/U$ that determines the relative strength of the single-particle and interaction terms.
@@
 \end{align}
-By design, these quantities are non-negative and normalized to a maximal value of one, independent of the system size $N$ or spin dimension $n$.
+By design, these non-negative quantities are normalized to a maximal value of one, independent of the system size $N$ or spin dimension $n$.
 In the remainder of this section we will assume that $n$ is even, both for the sake of experimental relevance (most relevant atomic nuclei are fermionic) and to avoid complications from parity effects.
+
+\subsection{Spin-polarized initial state}
 
@@
 \caption{
-Time-averaged magnetization $m_\MF$ and interaction energy $\bbk{\ss}_\MF$ for different spin dimensions $n$ (indicated in the legend) as determined by mean-field simulations of $N=100$ spins initially in the $x$-polarized state $\ket\x^{\otimes N}$ up to a time $T=10^5/U$.
-Insets show same data after normalizing $J\phi/U\to J\phi/U \times (n/2)^{1/3}$, as well as transforming $\bbk{\bar\sigma_\x}_\MF$ and $\bbk{\ss}_\MF$ according to Eq.~\eqref{eq:rescale}.
+Time-averaged magnetization $m_\MF$ and interaction energy $\bbk{\ss}_\MF$ for different spin dimensions $n$ (indicated in the legend) as determined by mean-field simulations of $N=100$ spins initially in the $x$-polarized state $\ket\x^{\otimes N}$ for a time $T=10^5/U$.
+Insets show same data after normalizing $J\phi/U\to J\phi/U \times (n/2)^{1/3}$, as well as transforming $m_\MF$ and $\bbk{\ss}_\MF$ according to Eq.~\eqref{eq:rescale}.
 }
@@
 As expected, the spin model exhibits a mean-field dynamical phase transition between an ordered phase at small $J\phi/U$ and a disordered phase at large $J\phi/U$.
-The ordered phase has a non-zero magnetization $m_\MF$ and an interaction energy $\bbk{\ss}_\MF$ that asymptotically approaches its maximal value as $J\phi/U\to0^+$.
-The disordered phase has no (time-averaged) magnetization, $m_\MF=0$, but an interaction energy $\bbk{\ss}_\MF$ that nonetheless indicates persistent nontrivial inter-spin correlations when $n>2$.
-These nontrivial inter-spin correlations vanish as $J\phi/U\to\infty$, in which case $\bbk{\ss}_\MF$ approaches the minimal value allowed by conservation laws (clarified below).
+The ordered phase has a non-zero magnetization $m_\MF$ and an interaction energy $\bbk{\ss}_\MF$ that asymptotically approach their maximal values as $J\phi/U\to0^+$.
+The disordered phase has no (time-averaged) magnetization, $m_\MF=0$, but the interaction energy $\bbk{\ss}_\MF$ nonetheless indicates persistent nontrivial inter-spin correlations when $n>2$.
+These nontrivial correlations vanish as $J\phi/U\to\infty$, in which case $\bbk{\ss}_\MF$ approaches the minimal value allowed by conservation laws (clarified below).
 By minimizing the reduced field $J\phi/U$ for which $m_\MF=0$, we numerically find that the transition between ordered and disordered phases occurs at a critical field $\p{J\phi/U}_{\t{crit}}=\p{n/2}^{-\alpha}$ with $\alpha\approx1/3$ (see Figure \ref{fig:crit_fields}).
@@
 \end{align}
-where we have used the fact that the axial magnetizations $\bk{s_{\z,q}}=0$ are conserved.
+where we have used the fact that the axial magnetizations $\bk{s_{\z,q}}=\frac1N\bk{S_\z}$ within the PS manifold, and $\bk{S_\z}=0$ is conserved by $H_{\t{spin}}^{\t{eff}}$.
 The weak-field effective Hamiltonian preserves permutational symmetry, so $\bbk{\ss}_\MF\to1$ as $J\phi/U\to0^+$.
-In addition, invariance of the initial state $\ket\x^{\otimes N}$ and the effective Hamiltonian $H_{\t{spin}}^{\t{eff}}$ under rotation by $e^{-\i\pi S_\x}$ implies that
-\begin{align}
-  \bk{S_\y} = \bk{e^{\i\pi S_\x} S_\y e^{-\i\pi S_\x}} = -\bk{S_\y}
-\end{align}
-at all times, so $\bk{S_\y}=0$.
-Together with permutational symmetry, the fact that $\bk{S_\y}=\bk{S_\z}=0$ implies that the magnetization $m_\MF$ is determined by the time-average of $\sigma_\x\equiv s_\x/s$ for a single spin:
+Moreover, permutational symmetry together with the fact that $\bk{S_\z}=\bk{S_\y}=0$ are conserved\footnote{Both the initial state $\ket\x^{\otimes N}$ and the spin Hamiltonian $H_{\t{spin}}$ are invariant under the action of $U_\x=e^{-\i\pi S_\x}\sum_{q>0}\v s_q\c\v s_{-q}$, so $\bk{S_\y}=\bk{U_\x^\dag S_\y U_\x}=-\bk{S_\y}=0$ at all times.} implies that the magnetization $m_\MF$ is determined by the time-average of $\sigma_\x\equiv s_\x/s$ for a single spin:
 \begin{align}
@@
   \lim_{J\phi/U\to0^+} m_\MF
-  = \abs{\f1{2\pi} \int_0^{2\pi} \dd\tau \p{\cos\tau}^{n-2}}
+  = \f1{2\pi} \int_0^{2\pi} \dd\tau \p{\cos\tau}^{n-2}
   = \gamma\p{\f{n}{2}}.
 \end{align}
+
+\subsection{Multi-cat initial states}
+
+We now consider the same setup as above, but with the initial ``multi-cat'' states $\ket{\x\x}^{\otimes N}$ and $\ket{\x\x_\i}^{\otimes N}$, where
+\begin{align}
+  \ket{\x\x} &\equiv \f{\ket{\x} + \ket{-\x}}{\sqrt{2}}
+  \simeq e^{-\i\frac{\pi}{2}s_\y^2} \ket{-s}, \\
+  \ket{\x\x_\i} &\equiv \f{\ket{\x} + \p{-1}^s \ket{-\x}}{\sqrt{2}}
+  \simeq e^{-\i\frac{\pi}{2}s_\x} \ket{\x\x}.
+\end{align}
+Here $\ket{-\x}\equiv e^{-\i\frac{\pi}{2}s_\y}\ket{-s}$, and $\simeq$ denotes equality up to a global phase.
 

### 2021-04-05 sun_phases/scripts/compute_boson_states.py

@@ -3,16 +3,15 @@
 import os, sys, time, scipy
 import numpy as np
 
-from dicke_methods import spin_op_z_dicke, spin_op_y_dicke, coherent_spin_state
-from boson_methods import polarized_state, boson_mft_state, evolve_mft, \
-    compute_avg_var_vals
+from dicke_methods import spin_op_vec_dicke, coherent_spin_state
+from boson_methods import boson_mft_state, evolve_mft, compute_avg_var_vals
 
 init_state_str = sys.argv[1]
 spin_dim = int(sys.argv[2])
 spin_num = int(sys.argv[3])
 log10_field = sys.argv[4]
 
-assert( init_state_str in [ "X", "XX" ] )
+assert( init_state_str in [ "X", "XX", "XXI" ] )
 assert( spin_dim % 2 == 0 )
 assert( spin_num % 2 == 0 )
 
@@ -32,24 +31,29 @@ def data_file(tag):
 genesis = time.time()
 ####################
 
+sz, sx, sy = spin_op_vec_dicke(spin_dim-1)
+
 # construct inhomogeneous magnetic field
 field = 10**float(log10_field)
-sz = spin_op_z_dicke(spin_dim-1).diagonal()
 def spin_coeff(qq):
     spin_angle = 2*np.pi*(qq+1/2)/spin_num
     return np.sin(spin_angle)
-field_data = [ sz, field * spin_coeff(np.arange(spin_num)) ]
+field_data = [ sz.diagonal(), field * spin_coeff(np.arange(spin_num)) ]
 
-# construct initial state
+# construct initial spin state
+state_x = coherent_spin_state("+X", spin_dim-1)
+state_dn = coherent_spin_state("-Z", spin_dim-1)
+state_xx = scipy.sparse.linalg.expm_multiply(-1j * np.pi/2 * sy @ sy, state_dn)
+state_xxi = scipy.sparse.linalg.expm_multiply(-1j * np.pi/2 * sx, state_xx)
 if init_state_str == "X":
-    init_state = polarized_state("+X", spin_dim, spin_num)
+    init_state = state_x
 elif init_state_str == "XX":
-    sy = spin_op_y_dicke(spin_dim-1)
-    state_z = coherent_spin_state("+Z", spin_dim-1)
-    state_xx = scipy.sparse.linalg.expm_multiply(-1j * np.pi/2 * sy @ sy, state_z)
-    init_state = boson_mft_state(state_xx, spin_dim, spin_num)
+    init_state = state_xx
[…]

### 2021-04-05 sun_phases/scripts/plot_boson_states.py

@@ -14,7 +14,7 @@ from boson_methods import extract_avg_var_state
 init_state_str = sys.argv[1]
 spin_num = int(sys.argv[2])
 
-assert( init_state_str in [ "X", "XX" ] )
+assert( init_state_str in [ "X", "XX", "XXI" ] )
 
 data_dir = "../data/spin_bosons/"
 fig_dir = "../figures/spin_bosons/"
@@ -108,10 +108,7 @@ for dim_idx, dim in enumerate(dims):
     plt.savefig(fig_dir + f"var_spect_{init_state_str}_n{dim:02d}.pdf")
     plt.close()
 
-    # skip plotting m_MF and <<ss>>_MF for XX initial state
-    if init_state_str == "XX": continue
-
-    # plot main data
+    # plot magnetization and interaction data
     marker, color = marker_color[dim]
     kwargs = dict( color = color, label = dim, zorder = -dim )
     axes[0].semilogx(fields, mean_mag, marker, **kwargs)
@@ -124,6 +121,9 @@ for dim_idx, dim in enumerate(dims):
     sub_axes[0].semilogx(fields_reduced, mean_mag_inset, marker, **kwargs)
     sub_axes[1].semilogx(fields_reduced, mean_int_inset, marker, **kwargs)
 
+    # only locate phase transition for x-polarized initial state
+    if init_state_str != "X": continue
+
     # locate when m_MF first hits zero
     dx = mean_mag[1:] - mean_mag[:-1]
     dh = log10_fields[1:] - log10_fields[:-1]
@@ -135,8 +135,6 @@ for dim_idx, dim in enumerate(dims):
     indices = slice(zero_start-3, zero_start)
     fit = np.polyfit(log10_fields[indices], mean_mag[indices], 2)
     crits[dim_idx] = 10**min(np.roots(fit)[-1], log10_fields[zero_start])
-
-if init_state_str == "XX": exit()
 
 # label axes and set axis ticks
 axes[0].set_ylabel(r"$m_\MF$")
@@ -161,6 +159,8 @@ plt.subplots_adjust(hspace = 0.08)
 kwargs = dict( bbox_extra_artists = (legend,), bbox_inches = "tight" )
 plt.savefig(fig_dir + f"mean_mag_int.pdf", **kwargs)
 plt.close()
+
+if init_state_str != "X": exit()
 
 ##################################################
 

### 2021-04-07 sun_phases/spin_model_paper/main.tex
added Lax vector analysis (that might be removed later...)

@@
 \centering
-\includegraphics{mean_mag_int.pdf}
+\includegraphics{mean_mag_int_X.pdf}
 \caption{
@@
 }
-\label{fig:mean_mag_int}
+\label{fig:mean_mag_int_X}
 \end{figure}
 
-Figure \ref{fig:mean_mag_int} shows the time-averages of the magnetization $m_\MF$ and interaction energy $\bbk{\ss}_\MF$ as computed by mean-field simulations of $N=100$ spins in an initial $x$-polarized state $\ket\x^{\otimes N}$, where
+Figure \ref{fig:mean_mag_int_X} shows the time-averages of the magnetization $m_\MF$ and interaction energy $\bbk{\ss}_\MF$ as computed by mean-field simulations of $N=100$ spins initially in the $x$-polarized state $\ket\x^{\otimes N}$, where
 \begin{align}
@@
   = \f1{2^s} \sum_\mu { 2s \choose s+\mu }^{1/2} \ket{\mu}.
+  \label{eq:state_x}
 \end{align}
@@
 These nontrivial correlations vanish as $J\phi/U\to\infty$, in which case $\bbk{\ss}_\MF$ approaches the minimal value allowed by conservation laws (clarified below).
-By minimizing the reduced field $J\phi/U$ for which $m_\MF=0$, we numerically find that the transition between ordered and disordered phases occurs at a critical field $\p{J\phi/U}_{\t{crit}}=\p{n/2}^{-\alpha}$ with $\alpha\approx1/3$ (see Figure \ref{fig:crit_fields}).
+By minimizing the reduced field $J\phi/U$ for which $m_\MF=0$, we numerically find that the transition between ordered and disordered phases occurs at a critical field $\p{J\phi/U}_{\t{crit}}=\p{n/2}^{-\alpha}$ with $\alpha\approx1/3$ (see Figure \ref{fig:crit_fields_X}).
 
@@
 \centering
-\includegraphics{crit_fields.pdf}
+\includegraphics{crit_fields_X.pdf}
 \caption{
@@
 }
-\label{fig:crit_fields}
+\label{fig:crit_fields_X}
 \end{figure}
 
-As shown in insets of Figure \ref{fig:mean_mag_int}, mean-field results for different spin dimensions $n$ collapse onto each other when normalizing the field $J\phi/U$ to its critical value, $\p{J\phi/U}\to\p{J\phi/U}\times\p{n/2}^{1/3}$, and rescaling
+As shown in insets of Figure \ref{fig:mean_mag_int_X}, mean-field results for different spin dimensions $n$ collapse onto each other when normalizing the field $J\phi/U$ to its critical value, $\p{J\phi/U}\to\p{J\phi/U}\times\p{n/2}^{1/3}$, and rescaling
 \begin{align}
@@
 \end{align}
-where we have used the fact that the axial magnetizations $\bk{s_{\z,q}}=\frac1N\bk{S_\z}$ within the PS manifold, and $\bk{S_\z}=0$ is conserved by $H_{\t{spin}}^{\t{eff}}$.
+where we have used the fact that the axial magnetizations $\bk{s_{\z,q}}=\frac1N\bk{S_\z}$ within the PS manifold, and $\bk{S_\z}=0$ is conserved by $H_{\t{spin}}$.
 The weak-field effective Hamiltonian preserves permutational symmetry, so $\bbk{\ss}_\MF\to1$ as $J\phi/U\to0^+$.
-Moreover, permutational symmetry together with the fact that $\bk{S_\z}=\bk{S_\y}=0$ are conserved\footnote{Both the initial state $\ket\x^{\otimes N}$ and the spin Hamiltonian $H_{\t{spin}}$ are invariant under the action of $U_\x=e^{-\i\pi S_\x}\sum_{q>0}\v s_q\c\v s_{-q}$, so $\bk{S_\y}=\bk{U_\x^\dag S_\y U_\x}=-\bk{S_\y}=0$ at all times.} implies that the magnetization $m_\MF$ is determined by the time-average of $\sigma_\x\equiv s_\x/s$ for a single spin:
+Moreover, both the spin Hamiltonian $H_{\t{spin}}$ and the initial state $\ket\x^{\otimes N}$ are invariant (up to global phase) under the action of $R_\x = e^{-\i\pi S_\x}\sum_{q>0}\v s_q\c\v s_{-q}$, so
+\begin{align}
+  \bk{S_\y} = \bk{R_\x S_\y R_\x^\dag} = -\bk{S_\y} = 0
+\end{align}
+at all times.
+Conservation of $\bk{S_\z}=\bk{S_\y}=0$ and permutational symmetry together imply that the magnetization $m_\MF$ is determined by the time-average of $\sigma_\x\equiv s_\x/s$ for a single spin:
 \begin{align}
@@
 \end{align}
-Here $\ket{-\x}\equiv e^{-\i\frac{\pi}{2}s_\y}\ket{-s}$, and $\simeq$ denotes equality up to a global phase.
+Here $\simeq$ denotes equality up to a global phase, and $\ket{-\x}$ is a state polarized along $-x$, defined similarly to $\ket\x$ in Eq.~\eqref{eq:state_x}:
+\begin{align}
+  \ket{-\x} \equiv e^{-\i\frac{\pi}{2}s_\y}\ket{-s}
+  = \f1{2^s} \sum_\mu \p{-1}^{s+\mu}
+  { 2s \choose s+\mu }^{1/2} \ket{\mu}.
+\end{align}
 
-\vspace{3cm}
+\begin{figure}
+\centering
+\includegraphics{mean_mag_int_XX.pdf}
+\caption{
+Numerical results similar to Figure \ref{fig:mean_mag_int_X}, but for an initial multi-cat state $\ket{\x\x}^{\otimes N}$:
+time-averaged magnetization $m_\MF$ and interaction energy $\bbk{\ss}_\MF$ for different spin dimensions $n$ (indicated in the legend) as determined by mean-field simulations of $N=100$ spins for a time $T=10^5/U$.
+}
+\label{fig:mean_mag_int_XX}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{mean_mag_int_XXI.pdf}
+\caption{
+Numerical results similar to Figure \ref{fig:mean_mag_int_X}, but for an initial multi-cat state $\ket{\x\x_\i}^{\otimes N}$:
+time-averaged magnetization $m_\MF$ and interaction energy $\bbk{\ss}_\MF$ for different spin dimensions $n$ (indicated in the legend) as determined by mean-field simulations of $N=100$ spins for a time $T=10^5/U$.
+Insets show same data after normalizing $J\phi/U\to J\phi/U \times (n/2)^{1/3}$, as well as transforming $m_\MF$ and $\bbk{\ss}_\MF$ according to Eq.~\eqref{eq:rescale}.
+}
+\label{fig:mean_mag_int_XXI}
+\end{figure}
 
@@
 \begin{align}
-  \i \d_t b_{\mu q}
+  \i \partial_t b_{\mu q}
   = -\f{U}{N} \sum_{\nu,p} b_{\nu p}^\dag b_{\mu p} b_{\nu q}
@@
   H = \sum_{\substack{\mu,\nu,\rho,\sigma\\j<k}}
-  h^{\mu\nu j}_{\rho\sigma k} s_{\mu\nu j} s_{\rho\sigma k}
+  g^{\mu\nu j}_{\rho\sigma k} s_{\mu\nu j} s_{\rho\sigma k}
   + \sum_{\mu,\nu,j} \epsilon_{\mu\nu j} s_{\mu\nu j},
@@
 \end{align}
-where $\mu,\nu$ index orthogonal states of an $n$-level spin; $j,k$ index one of $N$ spins; $h^{\mu\nu j}_{\rho\sigma k}$ and $\epsilon_{\mu\nu j}$ are scalars; and $s_{\mu\nu j}=\op{\mu}{\nu}_j$ is a transition operator for spin $j$.
-Strictly speaking, Eq.~\eqref{eq:quadratic_spin} only defines the couplings $h^{\mu\nu j}_{\rho\sigma k}$ for $j<k$, so we enforce $h^{\mu\nu k}_{\rho\sigma j}=h^{\mu\nu j}_{\rho\sigma k}$ and $h^{\mu\nu j}_{\rho\sigma j}=0$ for completion.
+where $\mu,\nu$ index orthogonal states of an $n$-level spin; $j,k$ index one of $N$ spins; $g^{\mu\nu j}_{\rho\sigma k}$ and $\epsilon_{\mu\nu j}$ are scalars; and $s_{\mu\nu j}=\op{\mu}{\nu}_j$ is a transition operator for spin $j$.
+Strictly speaking, Eq.~\eqref{eq:quadratic_spin} only defines the couplings $g^{\mu\nu j}_{\rho\sigma k}$ for $j<k$, so we enforce $g^{\mu\nu k}_{\rho\sigma j}=g^{\mu\nu j}_{\rho\sigma k}$ and $g^{\mu\nu j}_{\rho\sigma j}=0$ for completion.
 Decomposing spin operators into Schwinger bosons as $s_{\mu\nu j}=b_{\mu j}^\dag b_{\nu j}$, where $b_{\nu j}$ a annihilates a boson of type $\nu$ on site $j$, we can write this Hamiltonian as
@@
   H = \sum_{\substack{\mu,\nu,\rho,\sigma\\j<k}}
-  h^{\mu\nu j}_{\rho\sigma k}
+  g^{\mu\nu j}_{\rho\sigma k}
   b_{\mu j}^\dag b_{\nu j} b_{\rho k}^\dag b_{\sigma k}
@@
 \begin{align}
-  \i \d_t b_{\alpha\ell} = \sp{b_{\alpha\ell}, H}
+  \i \partial_t b_{\alpha\ell} = \sp{b_{\alpha\ell}, H}
   &= \sum_{\substack{\mu,\nu,\rho,\sigma\\j<k}}
-  h^{\mu\nu j}_{\rho\sigma k}
+  g^{\mu\nu j}_{\rho\sigma k}
   \sp{b_{\alpha\ell}, b_{\mu j}^\dag b_{\nu j} b_{\rho k}^\dag b_{\sigma k}}
@@
   \sp{b_{\alpha\ell}, b_{\mu j}^\dag b_{\nu j}} \\
-  &= \sum_{\mu,\nu,\rho,\sigma,k} h^{\mu\nu\ell}_{\rho\sigma k}
+  &= \sum_{\mu,\nu,\rho,\sigma,k} g^{\mu\nu\ell}_{\rho\sigma k}
   \sp{b_{\alpha\ell}, b_{\mu\ell}^\dag b_{\nu\ell}}
@@
   &= \sum_{\mu,\nu} \p{\sum_{\rho,\sigma,k}
-    h^{\mu\nu\ell}_{\rho\sigma k} b_{\rho k}^\dag b_{\sigma k}
+    g^{\mu\nu\ell}_{\rho\sigma k} b_{\rho k}^\dag b_{\sigma k}
     + \epsilon_{\mu\nu\ell}}
@@
 \begin{align}
-  \i \d_t b_{\alpha\ell}
+  \i \partial_t b_{\alpha\ell}
   = \sum_\nu \p{\sum_{\rho,\sigma,k}
-    h^{\alpha\nu\ell}_{\rho\sigma k} b_{\rho k}^\dag b_{\sigma k}
+    g^{\alpha\nu\ell}_{\rho\sigma k} b_{\rho k}^\dag b_{\sigma k}
     + \epsilon_{\alpha\nu\ell}} b_{\nu\ell}.
 \end{align}
-In the case of uniform SU($n$)-symmetric interactions of the form $\frac{h}{2}\v S\c\v S$ and a diagonal external field, we have
+In the case of uniform SU($n$)-symmetric interactions of the form $\frac{g}{2}\v S\c\v S$ and a diagonal external field, we have
 \begin{align}
-  h^{\alpha\nu\ell}_{\rho\sigma k}
-  = h \times \delta_{\alpha\sigma} \delta_{\nu\rho},
+  g^{\alpha\nu\ell}_{\rho\sigma k}
+  = g \times \delta_{\alpha\sigma} \delta_{\nu\rho},
   &&
@@
 \begin{align}
-  \i \d_t b_{\alpha\ell}
-  = h \sum_{\nu,k} b_{\nu k}^\dag b_{\alpha k} b_{\nu\ell}
+  \i \partial_t b_{\alpha\ell}
+  = g \sum_{\nu,k} b_{\nu k}^\dag b_{\alpha k} b_{\nu\ell}
   + \epsilon_{\alpha\ell} b_{\alpha\ell}.
 \end{align}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Lax vector analysis}
+
+We start with the spin Hamiltonian
+\begin{align}
+  H_{\t{spin}}
+  = -\f{U}{2N} \sum_{\mu,\nu} S_{\mu\nu} S_{\nu\mu}
+  + J\phi \sum_q \sin\p{q} s_{\z,q},
+\end{align}
+where $S_{\mu\nu} = \sum_q s_{\mu\nu q}$.
+The single-body operators involved have normalizations
+\begin{align}
+  \tr\p{s_{\mu\nu q}^\dag s_{\mu\nu q}} = 1
+  &&
+  \t{and}
+  &&
+  \tr\p{s_{\z,q}^\dag s_{\z,q}}
+  = \sum_\mu \mu^2
+  = \f1{12} (n+1) n (n-1)
+  \equiv \eta^2.
+\end{align}
+The Lax formulation requires all single-body operators involved to have the same normalization, so we substitute $s_{\tilde\z,q}\equiv s_{\z,q}/\eta$ to expand
+\begin{align}
+  U^{-1} H_{\t{spin}}
+  = -\f1{2N} \sum_{\mu,\nu} S_{\mu\nu} S_{\nu\mu}
+  + \eta h \sum_q \sin\p{q} s_{\tilde\z,q},
+  &&
+  \t{where}
+  &&
+  h \equiv \f{J\phi}{U}.
+\end{align}
+In the thermodynamic limit $N\to\infty$, the ``normalized'' Lax vector $\vec\ell\p{z}$ associated with $H_{\t{spin}}$ has components
+\begin{align}
+  \ell_\alpha\p{z}
+  = \lim_{N\to\infty} \f1N \sum_q \f{s_{\alpha,q}}{z-\sin q}
+    + \delta_{\alpha,\tilde\z} \, \eta h,
+\end{align}
+where $\alpha$ indexes elements of basis $\set{s_\alpha}$ of self-adjoint generators of SU($n$), with normalization $\tr\p{s_\alpha^2}=1$.
+Within the permutationally symmetric manifold, we can replace $s_{\alpha,q}\to\bar s_\alpha\equiv\frac1N\sum_q s_{\alpha,q}$ at the cost of $O(1/N)$ errors that vanish as $N\to\infty$, so
+\begin{align}
+  \ell_\alpha\p{z}
+  = \I\p{z} \bar s_\alpha
+  + \delta_{\alpha,\tilde\z} \, \eta h,
+\end{align}
+where
+\begin{align}
+  \I\p{z} \equiv \lim_{N\to\infty} \f1N \sum_q \f1{z-\sin\p{q}}
+  = \f1{2\pi} \int_{-\pi}^\pi \f{\dd q}{z-\sin\p{q}}
+  = \f1{\sqrt{z^2-1}} \times
+  \begin{cases}
+    +1 & \arg\p{z} \in (-\pi/2,\pi/2] \\
+    -1 & \t{otherwise}
+  \end{cases}
+  &&
+  \t{for}
+  &&
+  z \notin \sp{-1,1}.
+\end{align}
+The ``squared norm'' of the Lax vector is therefore
+\begin{align}
+  \vec\ell\p{z}^2
+  = \sum_\alpha \ell_\alpha\p{z}^2
+  = \I\p{z}^2 \sum_{\alpha\ne\tilde\z} \bar s_\alpha^2
+  + \sp{\I\p{z} \bar s_{\tilde\z} + \eta h}^2,
+\end{align}
+where we can define $Q^2\equiv\sum_\alpha \bar s_\alpha^2$ to simplify
+\begin{align}
+  \vec\ell\p{z}^2
+  = \I\p{z}^2 \p{Q^2 - \bar s_{\tilde\z}^2}
+  + \sp{\I\p{z} \bar s_{\tilde\z} + \eta h}^2
+  = \I\p{z}^2 Q^2 + \eta^2 h^2
+  + 2 \I\p{z} \eta h \bar s_{\tilde\z}.
+\end{align}
+For initial states with $\bar s_\z=0$, we thus find that
+\begin{align}
+  \vec\ell\p{z}^2 = \f{Q^2}{z^2-1} + \eta^2 h^2,
+\end{align}
+which is zero when
+\begin{align}
+  z = \pm \sqrt{1 - \p{\f{Q}{\eta h}}^2}.
+\end{align}
+These roots change character when $z=0$, suggesting that the critical field $h_{\t{crit}}$ separating dynamical phases satisfies
+\begin{align}
+  h_{\t{crit}}^2 = \f{Q^2}{\eta^2}.
+\end{align}
+For a permutationally symmetric state, up to vanishing $O(1/N)$ corrections we can expand
+\begin{align}
+  Q^2 = \sum_\alpha \bar s_\alpha^2
+  = \sum_{\mu,\nu} \bar s_{\mu\nu} \bar s_{\nu\mu} - \f1n
+  = 1 - \f1n
+  = \f{n-1}{n},
+\end{align}
+which implies that
+\begin{align}
+  h_{\t{crit}}^2 = \f{n-1}{n} \times \f{12}{(n+1)n(n-1)}
+  = \f{12}{n^2\p{n+1}}.
+\end{align}
+This Lax analysis correctly predicts that $h_{\t{crit}}=1$ when $n=2$, but otherwise predicts $h_{\t{crit}}\sim n^{-3/2}$, which is inconsistent with the finding that $h_{\t{crit}}\sim n^{-1/3}$ in mean-field simulations (see Figure \ref{fig:crit_fields_X} of the main text).
 

### 2021-04-07 sun_phases/spin_model_paper/main.tex
added discussion of XX MFT results

@@
 
+\newcommand{\xx}{\x\x}
+\newcommand{\xxi}{\x\x_\i}
+\newcommand{\X}{\text{X}}
+\newcommand{\XX}{\X\X}
+\newcommand{\XXI}{\X\X_\i}
 \newcommand{\su}{\mathfrak{su}}
@@
 \begin{document}
+\interfootnotelinepenalty=10000 % don't split footnotes between pages
 
@@
 \caption{
-Time-averaged magnetization $m_\MF$ and interaction energy $\bbk{\ss}_\MF$ for different spin dimensions $n$ (indicated in the legend) as determined by mean-field simulations of $N=100$ spins initially in the $x$-polarized state $\ket\x^{\otimes N}$ for a time $T=10^5/U$.
+Time-averaged magnetization $m_\MF$ and interaction energy $\bbk{\ss}_\MF$ for different spin dimensions $n$ (indicated in the legend) as determined by mean-field simulations of $N=100$ spins initially in the $x$-polarized state $\ket\X$ for a time $T=10^5/U$.
 Insets show same data after normalizing $J\phi/U\to J\phi/U \times (n/2)^{1/3}$, as well as transforming $m_\MF$ and $\bbk{\ss}_\MF$ according to Eq.~\eqref{eq:rescale}.
@@
 
-Figure \ref{fig:mean_mag_int_X} shows the time-averages of the magnetization $m_\MF$ and interaction energy $\bbk{\ss}_\MF$ as computed by mean-field simulations of $N=100$ spins initially in the $x$-polarized state $\ket\x^{\otimes N}$, where
+Figure \ref{fig:mean_mag_int_X} shows the time-averages of the magnetization $m_\MF$ and interaction energy $\bbk{\ss}_\MF$ as computed by mean-field simulations of $N=100$ spins initially in the $x$-polarized state $\X\equiv\ket\x^{\otimes N}$, where
 \begin{align}
@@
 \caption{
-The critical value of $\p{J\phi/U}_{\t{crit}}$ as determined by mean-field simulations of $N=100$ spins initially in the $x$-polarized state $\ket\x^{\otimes N}$.
+The critical value of $\p{J\phi/U}_{\t{crit}}$ as determined by mean-field simulations of $N=100$ spins initially in the $x$-polarized state $\ket\X$.
 A single-parameter fit to $\p{J\phi/U}_{\t{crit}}=\p{n/2}^{-\alpha}$ finds $\alpha=0.333(5)$, and $\alpha=1/3$ is consistent with all mean-field results to within an uncertainty determined by the resolution of $J\phi/U$ in mean-field simulations.
@@
   = \gamma\p{n}.
+  \label{eq:lim_ss}
 \end{align}
@@
 The weak-field effective Hamiltonian preserves permutational symmetry, so $\bbk{\ss}_\MF\to1$ as $J\phi/U\to0^+$.
-Moreover, both the spin Hamiltonian $H_{\t{spin}}$ and the initial state $\ket\x^{\otimes N}$ are invariant (up to global phase) under the action of $R_\x = e^{-\i\pi S_\x}\sum_{q>0}\v s_q\c\v s_{-q}$, so
+Moreover, both the spin Hamiltonian $H_{\t{spin}}$ and the initial state $\ket\X$ are invariant (up to global phase) under the action of $R_\x \equiv e^{-\i\pi S_\x}\sum_{q>0}\v s_q\c\v s_{-q}$ (i.e.~a $\pi$ rotation about the $x$ axis followed by a permutation of spins $q\leftrightarrow -q$), so
 \begin{align}
   \bk{S_\y} = \bk{R_\x S_\y R_\x^\dag} = -\bk{S_\y} = 0
+  \label{eq:sym_y}
 \end{align}
@@
 
-We now consider the same setup as above, but with the initial ``multi-cat'' states $\ket{\x\x}^{\otimes N}$ and $\ket{\x\x_\i}^{\otimes N}$, where
+We now consider the same setup as above, but with the initial ``multi-cat'' states $\ket\XX\equiv\ket\xx^{\otimes N}$ and $\ket\XXI\equiv\ket\xxi^{\otimes N}$, where
 \begin{align}
-  \ket{\x\x} &\equiv \f{\ket{\x} + \ket{-\x}}{\sqrt{2}}
+  \ket\xx &\equiv \f{\ket{\x} + \ket{-\x}}{\sqrt{2}}
   \simeq e^{-\i\frac{\pi}{2}s_\y^2} \ket{-s}, \\
-  \ket{\x\x_\i} &\equiv \f{\ket{\x} + \p{-1}^s \ket{-\x}}{\sqrt{2}}
-  \simeq e^{-\i\frac{\pi}{2}s_\x} \ket{\x\x}.
+  \ket\xxi &\equiv \f{\ket{\x} + \p{-1}^s \ket{-\x}}{\sqrt{2}}
+  \simeq e^{-\i\frac{\pi}{2}s_\x} \ket\xx.
 \end{align}
@@
 \caption{
-Numerical results similar to Figure \ref{fig:mean_mag_int_X}, but for an initial multi-cat state $\ket{\x\x}^{\otimes N}$:
-time-averaged magnetization $m_\MF$ and interaction energy $\bbk{\ss}_\MF$ for different spin dimensions $n$ (indicated in the legend) as determined by mean-field simulations of $N=100$ spins for a time $T=10^5/U$.
+A near-replica of Figure \ref{fig:mean_mag_int_X} for the initial state $\ket\XX$.
+The inset for interaction energy $\bbk{\ss}_\MF$ in Figure \ref{fig:mean_mag_int_X} subtracts off the minimal value of $\bbk{\ss}_\MF$ and rescales to lie on the interval $[0,1]$, as prescribed by Eq.~\eqref{eq:rescale}.
+Here the subtracting and rescaling is identical, but with a minimal value of $2\gamma\p{n}$ rather than $\gamma\p{n}$.
 }
@@
 \caption{
-Numerical results similar to Figure \ref{fig:mean_mag_int_X}, but for an initial multi-cat state $\ket{\x\x_\i}^{\otimes N}$:
-time-averaged magnetization $m_\MF$ and interaction energy $\bbk{\ss}_\MF$ for different spin dimensions $n$ (indicated in the legend) as determined by mean-field simulations of $N=100$ spins for a time $T=10^5/U$.
-Insets show same data after normalizing $J\phi/U\to J\phi/U \times (n/2)^{1/3}$, as well as transforming $m_\MF$ and $\bbk{\ss}_\MF$ according to Eq.~\eqref{eq:rescale}.
+A replica of Figure \ref{fig:mean_mag_int_X} for the initial state $\ket\XXI$.
+Insets show the same data shifted and rescaled identically to Figure \ref{fig:mean_mag_int_X}.
 }
@@
 \end{figure}
+
+Similarly to Figure \ref{fig:mean_mag_int_X}, Figures \ref{fig:mean_mag_int_XX} and \ref{fig:mean_mag_int_XXI} respectively show the time-averaged magnetization $m_\MF$ and interaction energy $\bbk{\ss}_\MF$ throughout mean-field dynamics of the initial states $\ket\XX$ and $\ket\XXI$.
+These figures exclude the trivial case of spin dimension $n=2$, for which $\ket\xx=\ket{-s}$ is an eigenstate of $H_{\t{spin}}$ and $\ket\xxi=e^{-i\frac{\pi}{2}s_\z}\ket\x$ is still an equatorially-polarized state.
+The first and perhaps most interesting observation to make about Figures \ref{fig:mean_mag_int_XX} and \ref{fig:mean_mag_int_XXI} is that they are different, signifying the importance of intra-spin coherences for the dynamical behavior of multilevel spin models.
+
+Unlike the case of an initial spin-polarized state $\ket\X$, the mean-field dynamics of the initial multi-cat state $\ket\XX$ (Figure \ref{fig:mean_mag_int_XX}) exhibit no sharp transition between distinct dynamical phases: the time-averaged magnetization $m_\MF=0$ for all values of the field $J\phi/U$, and the interaction energy $\bbk{\ss}_\MF$ smoothly crosses over from a maximal value of 1 to a minimal value of $2\gamma\p{n}$.
+The minimal value of $\bbk{\ss}_\MF$ approached as $J\phi/U\to\infty$ can be explained using arguments identical to those above Eq.~\eqref{eq:lim_ss}, which now imply that
+\begin{align}
+  \lim_{J\phi/U\to\infty} \bbk{\ss}_\MF
+  = \Tr\sp{\p{\diag\op{\xx}}^2}
+  = 2\gamma\p{n}.
+\end{align}
+The vanishing magnetization $m_\MF=0$ in Figure \ref{fig:mean_mag_int_XX} is protected by symmetries of $H_{\t{spin}}$ and $\ket\XX$.
+For all initial states that we have considered, the value of $\bk{S_\z}=0$ is conserved by the spin Hamiltonian $H_{\t{spin}}$.
+Moreover, as discussed above Eq.~\eqref{eq:sym_y}, the invariance of $H_{\t{spin}}$ and all initial states under the action of $R_\x\equiv e^{-\i\pi S_\x}\sum_{q>0}\v s_q\c\v s_{-q}$ implies that $\bk{S_\y}=0$ at all times.
+The state $\ket\XX$ is additionally invariant under the action of $R_\y\equiv e^{-\i\pi S_\y}\sum_{q>0}\v s_q\c\v s_{-q}$, which analogously ensures that $\bk{S_\x}=0$, and thereby $m_\MF=0$.
+
+Turning now to mean-field results for the initial multi-cat state $\ket\XXI$ in Figure \ref{fig:mean_mag_int_XXI}, we remark that the magnetization $m_\MF$ and interaction energy $\bbk{\ss}_\MF$ behave identically to those for the initial spin-polarized state $\ket\X$ in Figure \ref{fig:mean_mag_int_X}.
 
@@
 \end{align}
-which is zero when
+which is zero when\footnote{Strictly speaking, the zeros in Eq.~\eqref{eq:lax_zeros} occur at values of $z$ at which $\I\p{z}$ is undefined.
+We avoid this issue by analytically continuing $\I\p{z}^2$ to the interval $z\in(-1,1)$.}
 \begin{align}
   z = \pm \sqrt{1 - \p{\f{Q}{\eta h}}^2}.
+  \label{eq:lax_zeros}
 \end{align}

### 2021-04-09 sun_phases/spin_model_paper/main.tex
change notation for magnetization, add more text to multi-cat state section

@@
   = \sum_{\mu,\nu} A_{\mu\nu}^\dag B_{\mu\nu},
+  \label{eq:dot_product}
 \end{align}
@@
 \begin{align}
-  m_\MF \equiv \abs{\bbk{\vec\sigma}_\MF},
+  \sigma_\MF \equiv \abs{\bbk{\vec\sigma}_\MF},
   &&
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Spin-polarized initial state}
@@
 \caption{
-Time-averaged magnetization $m_\MF$ and interaction energy $\bbk{\ss}_\MF$ for different spin dimensions $n$ (indicated in the legend) as determined by mean-field simulations of $N=100$ spins initially in the $x$-polarized state $\ket\X$ for a time $T=10^5/U$.
-Insets show same data after normalizing $J\phi/U\to J\phi/U \times (n/2)^{1/3}$, as well as transforming $m_\MF$ and $\bbk{\ss}_\MF$ according to Eq.~\eqref{eq:rescale}.
+Time-averaged magnetization $\sigma_\MF$ and interaction energy $\bbk{\ss}_\MF$ for different spin dimensions $n$ (indicated in the legend) as determined by mean-field simulations of $N=100$ spins initially in the $x$-polarized state $\ket\X$ for a time $T=10^5/U$.
+Insets show same data after rescaling $J\phi/U\to J\phi/U \times (n/2)^{1/3}$, and transforming vertical axes according to Eq.~\eqref{eq:rescale}.
 }
@@
 
-Figure \ref{fig:mean_mag_int_X} shows the time-averages of the magnetization $m_\MF$ and interaction energy $\bbk{\ss}_\MF$ as computed by mean-field simulations of $N=100$ spins initially in the $x$-polarized state $\X\equiv\ket\x^{\otimes N}$, where
+Figure \ref{fig:mean_mag_int_X} shows the time-averages of the magnetization $\sigma_\MF$ and interaction energy $\bbk{\ss}_\MF$ as computed by mean-field simulations of $N=100$ spins initially in the $x$-polarized state $\X\equiv\ket\x^{\otimes N}$, where
 \begin{align}
@@
 As expected, the spin model exhibits a mean-field dynamical phase transition between an ordered phase at small $J\phi/U$ and a disordered phase at large $J\phi/U$.
-The ordered phase has a non-zero magnetization $m_\MF$ and an interaction energy $\bbk{\ss}_\MF$ that asymptotically approach their maximal values as $J\phi/U\to0^+$.
-The disordered phase has no (time-averaged) magnetization, $m_\MF=0$, but the interaction energy $\bbk{\ss}_\MF$ nonetheless indicates persistent nontrivial inter-spin correlations when $n>2$.
+The ordered phase has a non-zero magnetization $\sigma_\MF$ and an interaction energy $\bbk{\ss}_\MF$ that asymptotically approach their maximal values as $J\phi/U\to0^+$.
+The disordered phase has no (time-averaged) magnetization, $\sigma_\MF=0$, but the interaction energy $\bbk{\ss}_\MF$ nonetheless indicates persistent nontrivial inter-spin correlations when $n>2$.
 These nontrivial correlations vanish as $J\phi/U\to\infty$, in which case $\bbk{\ss}_\MF$ approaches the minimal value allowed by conservation laws (clarified below).
-By minimizing the reduced field $J\phi/U$ for which $m_\MF=0$, we numerically find that the transition between ordered and disordered phases occurs at a critical field $\p{J\phi/U}_{\t{crit}}=\p{n/2}^{-\alpha}$ with $\alpha\approx1/3$ (see Figure \ref{fig:crit_fields_X}).
+By minimizing the reduced field $J\phi/U$ for which $\sigma_\MF=0$, we numerically find that the transition between ordered and disordered phases occurs at a critical field $\p{J\phi/U}_{\t{crit}}=\p{n/2}^{-\alpha}$ with $\alpha\approx1/3$ (see Figure \ref{fig:crit_fields_X}).
 
@@
 \begin{align}
-  m_\MF \to \f{m_\MF}{\gamma\p{n/2}},
+  \sigma_\MF \to \f{\sigma_\MF}{\gamma\p{n/2}},
   &&
@@
 In the strong-field limit $J\phi/U\to\infty$, we can ignore interactions and treat spins as though they simply precess at different rates.
-The time-averaged transverse magnetization $m_\MF$ then trivially vanishes as $J\phi/U\to\infty$.
+The time-averaged transverse magnetization $\sigma_\MF$ then trivially vanishes as $J\phi/U\to\infty$.
 The interaction energy $\bk{\ss}_\MF = \bk{\bar{\v s}}_\MF \c \bk{\bar{\v s}}_\MF$, meanwhile, has contributions from
@@
 The weak-field effective Hamiltonian preserves permutational symmetry, so $\bbk{\ss}_\MF\to1$ as $J\phi/U\to0^+$.
-Moreover, both the spin Hamiltonian $H_{\t{spin}}$ and the initial state $\ket\X$ are invariant (up to global phase) under the action of $R_\x \equiv e^{-\i\pi S_\x}\sum_{q>0}\v s_q\c\v s_{-q}$ (i.e.~a $\pi$ rotation about the $x$ axis followed by a permutation of spins $q\leftrightarrow -q$), so
+Moreover, both the spin Hamiltonian $H_{\t{spin}}$ and the initial state $\ket\X$ are invariant (up to global phase) under the action of $\Pi R_\x^\pi$, where $R_\x^\theta \equiv e^{-\i\theta S_\x}$ is a rotation about the $x$ axis by $\theta$ and $\Pi\equiv\sum_{q>0}\v s_q\c\v s_{-q}$ is a permutation of spins $q\leftrightarrow-q$, which is to say that
 \begin{align}
-  \bk{S_\y} = \bk{R_\x S_\y R_\x^\dag} = -\bk{S_\y} = 0
-  \label{eq:sym_y}
+  e^{-\i t H_{\t{spin}}} \ket\X
+  &\simeq \p{\Pi R_\x^\pi e^{-\i t H_{\t{spin}}} {R_\x^\pi}^\dag \Pi^\dag} \Pi R_\x^\pi \ket\X
+  \label{eq:sym_x_start} \\
+  &\simeq \Pi R_\x^\pi e^{-\i t H_{\t{spin}}} \ket\X,
+\end{align}
+so
+\begin{align}
+  \bk{S_\y}
+  = \bk{{R_\x^\pi}^\dag \Pi^\dag S_\y \Pi R_\x^\pi}
+  = -\bk{S_\y} = 0
+  \label{eq:sym_x_end}
 \end{align}
 at all times.
-Conservation of $\bk{S_\z}=\bk{S_\y}=0$ and permutational symmetry together imply that the magnetization $m_\MF$ is determined by the time-average of $\sigma_\x\equiv s_\x/s$ for a single spin:
+Conservation of $\bk{S_\y}=\bk{S_\z}=0$ and permutational symmetry together imply that the magnetization $\sigma_\MF$ is determined by the time-average of $\sigma_\x\equiv s_\x/s$ for a single spin:
 \begin{align}
-  \lim_{J\phi/U\to0^+} m_\MF
+  \lim_{J\phi/U\to0^+} \sigma_\MF
   = \abs{\lim_{T\to\infty} \f1T \int_0^T \dd\tau
@@
 \begin{align}
-  \lim_{J\phi/U\to0^+} m_\MF
+  \lim_{J\phi/U\to0^+} \sigma_\MF
   = \f1{2\pi} \int_0^{2\pi} \dd\tau \p{\cos\tau}^{n-2}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Multi-cat initial states}
-
-We now consider the same setup as above, but with the initial ``multi-cat'' states $\ket\XX\equiv\ket\xx^{\otimes N}$ and $\ket\XXI\equiv\ket\xxi^{\otimes N}$, where
-\begin{align}
-  \ket\xx &\equiv \f{\ket{\x} + \ket{-\x}}{\sqrt{2}}
-  \simeq e^{-\i\frac{\pi}{2}s_\y^2} \ket{-s}, \\
-  \ket\xxi &\equiv \f{\ket{\x} + \p{-1}^s \ket{-\x}}{\sqrt{2}}
-  \simeq e^{-\i\frac{\pi}{2}s_\x} \ket\xx.
-\end{align}
-Here $\simeq$ denotes equality up to a global phase, and $\ket{-\x}$ is a state polarized along $-x$, defined similarly to $\ket\x$ in Eq.~\eqref{eq:state_x}:
-\begin{align}
-  \ket{-\x} \equiv e^{-\i\frac{\pi}{2}s_\y}\ket{-s}
-  = \f1{2^s} \sum_\mu \p{-1}^{s+\mu}
-  { 2s \choose s+\mu }^{1/2} \ket{\mu}.
-\end{align}
 
@@
 \caption{
-A near-replica of Figure \ref{fig:mean_mag_int_X} for the initial state $\ket\XX$.
+A corollary of Figure \ref{fig:mean_mag_int_X} for the initial state $\ket\XX$.
 The inset for interaction energy $\bbk{\ss}_\MF$ in Figure \ref{fig:mean_mag_int_X} subtracts off the minimal value of $\bbk{\ss}_\MF$ and rescales to lie on the interval $[0,1]$, as prescribed by Eq.~\eqref{eq:rescale}.
@@
 \caption{
-A replica of Figure \ref{fig:mean_mag_int_X} for the initial state $\ket\XXI$.
+A corollary of Figure \ref{fig:mean_mag_int_X} for the initial state $\ket\XXI$.
 Insets show the same data shifted and rescaled identically to Figure \ref{fig:mean_mag_int_X}.
@@
 
-Similarly to Figure \ref{fig:mean_mag_int_X}, Figures \ref{fig:mean_mag_int_XX} and \ref{fig:mean_mag_int_XXI} respectively show the time-averaged magnetization $m_\MF$ and interaction energy $\bbk{\ss}_\MF$ throughout mean-field dynamics of the initial states $\ket\XX$ and $\ket\XXI$.
-These figures exclude the trivial case of spin dimension $n=2$, for which $\ket\xx=\ket{-s}$ is an eigenstate of $H_{\t{spin}}$ and $\ket\xxi=e^{-i\frac{\pi}{2}s_\z}\ket\x$ is still an equatorially-polarized state.
+We now consider the same setup as above, but with the initial ``multi-cat'' states $\ket\XX\equiv\ket\xx^{\otimes N}$ and $\ket\XXI\equiv\ket\xxi^{\otimes N}$, where
+\begin{align}
+  \ket\xx \equiv \f{\ket{\x} + \ket{-\x}}{\sqrt{2}},
+  &&
+  \ket\xxi \equiv \f{\ket{\x} + \p{-1}^s \ket{-\x}}{\sqrt{2}},
+\end{align}
+and $\ket{-\x}$ is a state polarized along $-x$, defined similarly to $\ket\x$ in Eq.~\eqref{eq:state_x}:
+\begin{align}
+  \ket{-\x} \equiv e^{-\i\frac{\pi}{2}s_\y}\ket{-s}
+  = \f1{2^s} \sum_\mu \p{-1}^{s+\mu}
+  { 2s \choose s+\mu }^{1/2} \ket{\mu}.
+\end{align}
+Similarly to Figure \ref{fig:mean_mag_int_X}, Figures \ref{fig:mean_mag_int_XX} and \ref{fig:mean_mag_int_XXI} respectively show the time-averaged magnetization $\sigma_\MF$ and interaction energy $\bbk{\ss}_\MF$ throughout mean-field dynamics of the initial states $\ket\XX$ and $\ket\XXI$.
+These figures exclude the trivial case of spin dimension $n=2$, for which $\ket\xx=\ket{-s}$ is an eigenstate of $H_{\t{spin}}$ and $\ket\xxi=e^{-\i\frac{\pi}{2}s_\z}\ket\x\equiv\ket\y$.
 The first and perhaps most interesting observation to make about Figures \ref{fig:mean_mag_int_XX} and \ref{fig:mean_mag_int_XXI} is that they are different, signifying the importance of intra-spin coherences for the dynamical behavior of multilevel spin models.
 
-Unlike the case of an initial spin-polarized state $\ket\X$, the mean-field dynamics of the initial multi-cat state $\ket\XX$ (Figure \ref{fig:mean_mag_int_XX}) exhibit no sharp transition between distinct dynamical phases: the time-averaged magnetization $m_\MF=0$ for all values of the field $J\phi/U$, and the interaction energy $\bbk{\ss}_\MF$ smoothly crosses over from a maximal value of 1 to a minimal value of $2\gamma\p{n}$.
-The minimal value of $\bbk{\ss}_\MF$ approached as $J\phi/U\to\infty$ can be explained using arguments identical to those above Eq.~\eqref{eq:lim_ss}, which now imply that
+Unlike the case of an initial spin-polarized state $\ket\X$, the mean-field dynamics of the initial multi-cat state $\ket\XX$ (Figure \ref{fig:mean_mag_int_XX}) exhibit no sharp transition between distinct dynamical phases: the time-averaged magnetization $\sigma_\MF=0$ for all values of the field $J\phi/U$, and the interaction energy $\bbk{\ss}_\MF$ smoothly crosses over from a maximal value of 1 to a minimal value of $2\gamma\p{n}$.
+The minimal value of $\bbk{\ss}_\MF$ approached as $J\phi/U\to\infty$ can be explained with arguments identical to those above Eq.~\eqref{eq:lim_ss}, which now imply that
 \begin{align}
@@
 \end{align}
-The vanishing magnetization $m_\MF=0$ in Figure \ref{fig:mean_mag_int_XX} is protected by symmetries of $H_{\t{spin}}$ and $\ket\XX$.
+The vanishing magnetization $\sigma_\MF=0$ in Figure \ref{fig:mean_mag_int_XX} is protected by symmetries of $H_{\t{spin}}$ and $\ket\XX$.
 For all initial states that we have considered, the value of $\bk{S_\z}=0$ is conserved by the spin Hamiltonian $H_{\t{spin}}$.
-Moreover, as discussed above Eq.~\eqref{eq:sym_y}, the invariance of $H_{\t{spin}}$ and all initial states under the action of $R_\x\equiv e^{-\i\pi S_\x}\sum_{q>0}\v s_q\c\v s_{-q}$ implies that $\bk{S_\y}=0$ at all times.
-The state $\ket\XX$ is additionally invariant under the action of $R_\y\equiv e^{-\i\pi S_\y}\sum_{q>0}\v s_q\c\v s_{-q}$, which analogously ensures that $\bk{S_\x}=0$, and thereby $m_\MF=0$.
+Moreover, both the spin Hamiltonian $H_{\t{spin}}$ and the state $\ket\XX$ are invariant (up to global phase) under the action of $R_\z^\pi$, where $R_\z^\theta\equiv e^{-\i\theta S_\z}$.
+Similarly to Eqs.~\eqref{eq:sym_x_start}--\eqref{eq:sym_x_end}, this symmetry implies that
+\begin{align}
+  &\bk{S_\x} = \bk{{R_\z^\pi}^\dag S_\x R_\z^\pi} = -\bk{S_\x} = 0, \\
+  &\bk{S_\y} = \bk{{R_\z^\pi}^\dag S_\y R_\z^\pi} = -\bk{S_\y} = 0
+\end{align}
+at all times, so altogether $\sigma_\MF=0$.
 
-Turning now to mean-field results for the initial multi-cat state $\ket\XXI$ in Figure \ref{fig:mean_mag_int_XXI}, we remark that the magnetization $m_\MF$ and interaction energy $\bbk{\ss}_\MF$ behave identically to those for the initial spin-polarized state $\ket\X$ in Figure \ref{fig:mean_mag_int_X}.
+Turning now to mean-field results for the initial multi-cat state $\ket\XXI$ in Figure \ref{fig:mean_mag_int_XXI}, we remark that the magnetization $\sigma_\MF$ and interaction energy $\bbk{\ss}_\MF$ behave identically to those for the initial spin-polarized state $\ket\X$ in Figure \ref{fig:mean_mag_int_X}.
+This finding can be understood through the fact that
+\begin{align}
+  \ket\XXI \simeq R_\z^{\pi/2} T_\z^{\pi/2} \ket\X,
+\end{align}
+where $T_\z^\theta\equiv e^{-\i\theta S_\z^2}$.
+The operators $R_\z^\theta$ and $T_\z^\theta$ are generated by axial fields that respect permutational symmetry, and therefore commute with the spin Hamiltonian $H_{\t{spin}}$, so
+\begin{align}
+  e^{-\i t H_{\t{spin}}} \ket\XXI
+  &\simeq e^{-\i t H_{\t{spin}}} R_\z^{\pi/2} T_\z^{\pi/2} \ket\X \\
+  &\simeq R_\z^{\pi/2} T_\z^{\pi/2} e^{-\i t H_{\t{spin}}} \ket\X.
+\end{align}
+In turn, expanding $\ss$ according to Eq.~\eqref{eq:dot_product} shows that
+\begin{align}
+  {T_\z^\theta}^\dag {R_\z^\theta}^\dag \, \ss \, R_\z^\theta T_\z^\theta
+  = \ss,
+\end{align}
+which implies that the interaction energy $\bk{\ss}$ throughout dynamics of the initial multi-cat state $\ket\XXI$ is the same as that of the initial spin-polarized state $\ket\X$.
+
+To make sense of why the magnetization $\sigma_\MF$ should behave identically in Figure \ref{fig:mean_mag_int_XXI} for $\ket\XXI$ as in Figure \ref{fig:mean_mag_int_X} for $\ket\X$, we can ignore the time-independent factors $R_\z^\theta$, which only rotate the mean spin vector $\bbk{\vec\sigma}_\MF$ without changing its magnitude.
+
+\red{[explain $\sigma_\MF$ is identical in Figures \ref{fig:mean_mag_int_X} and \ref{fig:mean_mag_int_XXI}]}
 
@@
 \end{align}
-In the thermodynamic limit $N\to\infty$, the ``normalized'' Lax vector $\vec\ell\p{z}$ associated with $H_{\t{spin}}$ has components
+The normalized Lax vector $\vec\ell\p{z}$ associated with $H_{\t{spin}}$ has components
 \begin{align}
   \ell_\alpha\p{z}
-  = \lim_{N\to\infty} \f1N \sum_q \f{s_{\alpha,q}}{z-\sin q}
+  = \f1N \sum_q \f{s_{\alpha,q}}{z-\sin q}
   + \delta_{\alpha,\tilde\z} \, \eta h,
@@
 where $\alpha$ indexes elements of basis $\set{s_\alpha}$ of self-adjoint generators of SU($n$), with normalization $\tr\p{s_\alpha^2}=1$.
-Within the permutationally symmetric manifold, we can replace $s_{\alpha,q}\to\bar s_\alpha\equiv\frac1N\sum_q s_{\alpha,q}$ at the cost of $O(1/N)$ errors that vanish as $N\to\infty$, so
+Within the permutationally symmetric manifold, we can replace $s_{\alpha,q}\to\bar s_\alpha\equiv\frac1N\sum_q s_{\alpha,q}$ at the cost of $O(1/N)$ errors that vanish as $N\to\infty$, so taking this limit we find
 \begin{align}
@@
 \begin{align}
-  \I\p{z} \equiv \lim_{N\to\infty} \f1N \sum_q \f1{z-\sin\p{q}}
-  = \f1{2\pi} \int_{-\pi}^\pi \f{\dd q}{z-\sin\p{q}}
-  = \f1{\sqrt{z^2-1}} \times
-  \begin{cases}
-    +1 & \arg\p{z} \in (-\pi/2,\pi/2] \\
-    -1 & \t{otherwise}
-  \end{cases}
+  \I\p{z} \equiv \f1N \sum_q \f1{z-\sin\p{q}}
+  = \f1{2\pi} \int_0^{2\pi} \f{\dd q}{z-\sin\p{q}}
+  = \f1{\sqrt{z^2-1}}
   &&

### 2021-04-09 sun_phases/scripts/plot_boson_states.py

@@ -135,7 +135,7 @@ for dim_idx, dim in enumerate(dims):
     crits[dim_idx] = 10**min(np.roots(fit)[-1], log10_fields[zero_start])
 
 # label axes and set axis ticks
-axes[0].set_ylabel(r"$m_\MF$")
+axes[0].set_ylabel(r"$\sigma_\MF$")
 axes[1].set_ylabel(r"$\bbk{\bar{\bm s}\cdot\bar{\bm s}}_\MF$")
 axes[1].set_xlabel(r"$J\phi/U$")
 axes[1].set_xlim(0.1,10)

### 2021-04-15 sun_phases/spin_model_paper/main.tex
'finalized' main body of SU(n) spin model paper (prior to review/feedback)

@@
 \newcommand{\C}{\mathcal{C}}
+\newcommand{\D}{\mathcal{D}}
+\newcommand{\E}{\mathcal{E}}
 \newcommand{\I}{\mathcal{I}}
@@
 \newcommand{\XXI}{\X\X_\i}
+\newcommand{\spin}{\text{spin}}
+\newcommand{\eff}{\text{eff}}
+\newcommand{\crit}{\text{crit}}
+\newcommand{\MF}{\text{MF}}
 \newcommand{\su}{\mathfrak{su}}
-\newcommand{\MF}{\text{MF}}
 \renewcommand{\ss}{\bar{\v s}\c\bar{\v s}}
 \DeclareMathOperator{\diag}{diag}
+
+\newcommand{\col}{\underline}
+\newcommand{\mean}{\overline}
 
@@
 If the interaction energy $U_0$ is smaller than the single-particle bandwidth $\sim J$, then the mode-changing collisions in $H_{\t{int}}$ become off-resonant, motivating the frozen-mode approximation $\set{k,p}=\set{\ell,q}$ (which we assume for the remainder of this work)\footnote{Note that the frozen-mode approximation neglects correlated momentum-hopping terms of the form $c_{\pi-p,\mu}^\dag c_{\pi-q,\mu} c_{p\nu}^\dag c_{q\nu}$, which conserve both momentum and energy.
-We defer a careful treatment of these terms to future work, noting only that they vanish on the manifold of permutationally-symmetric spin states with one atom per lattice site.}.
+We defer a careful treatment of these terms to future work, noting only that they vanish on the manifold of permutationally symmetric spin states with one atom per lattice site.}.
 The operator content of terms with $k=\ell$ and $p=q$ takes the form $c_{\ell\mu}^\dag c_{\ell\mu} c_{q\nu}^\dag c_{q\nu}$, so these terms contribute an overall energy $\frac12NU$ that can be neglected in the absence of coherent atom number fluctuations.
@@
 (b,c) When the interaction energy $U_0$ is smaller than the single-particle bandwidth $\sim J$, the frozen-mode approximation enables writing the interaction Hamiltonian as a spin model consisting of exchange terms $\v s_p\c\v s_q$, which swap the states two spins pinned to momentum modes $p,q$.
-(d) Interactions thereby open an energy gap $U=U_0\times N/L$ between the manifold of permutationally-symmetric states of $N$ spins, and the orthogonal complement of states that break spin-permutation symmetry.
+(d) Interactions thereby open an energy gap $U=U_0\times N/L$ between the manifold of permutationally symmetric states of $N$ spins, and the orthogonal complement of states that break spin-permutation symmetry.
 }
@@
 In this sense, $\v s_p\c\v s_q$ is analogous to the enforcement of SU(2) spin alignment by ferromagnetic interactions, which similarly assigns (distinct) definite energies to anti-symmetric spin-0 singlet $\ket{\up\dn}-\ket{\dn\up}$ and the symmetric spin-1 triplets $\set{\ket{\up\up},\ket{\dn\dn},\ket{\up\dn}+\ket{\dn\up}}$.
-By summing over all pair-wise exchange terms $\v s_p\c\v s_q$, the interaction Hamiltonian $H_{\t{int}}$ energetically enforces a permutational symmetry among all spins, opening an energy gap $U$ between the manifold of all permutationally-symmetric (PS) states and the orthogonal complement of excited (e.g.~spin-wave) states that break permutational symmetry.
+By summing over all pair-wise exchange terms $\v s_p\c\v s_q$, the interaction Hamiltonian $H_{\t{int}}$ energetically enforces a permutational symmetry among all spins, opening an energy gap $U$ between the manifold of all permutationally symmetric (PS) states and the orthogonal complement of excited (e.g.~spin-wave) states that break permutational symmetry.
 See Figure \ref{fig:spin_model} for a summary of this section thus far.
@@
 External fields or additional interactions that respect permutational symmetry can induce nontrivial dynamics within the PS manifold.
-Even if additional Hamiltonian terms break permutational symmetry, as long as the coupling to states outside of the PS manifold is small (compared to the energy gap $U$) these terms can be restricted to the PS manifold perturbatively (see Appendix \ref{sec:pert_theory}).
+Even if additional Hamiltonian terms break permutational symmetry, as long as the coupling to states outside of the PS manifold is small (compared to the energy gap $U$) these terms can be restricted to the PS manifold perturbatively (see Appendix \ref{sec:pert_theory}) \cite{bravyi2011schrieffer}.
 Simulating dynamics within the PS manifold requires calculating matrix elements $\bk{\ell|\O|m}$ of spin operators $\O$ with respect to PS states $\ket\ell,\ket{m}$; we discuss this calculation in Appendix \ref{sec:PS_ops}.
@@
 \begin{align}
-  H_{\t{spin}} = -\f{U}{2N} \v S\c\v S + J\phi \sum_q \sin\p{q} s_{\z,q},
+  H_\spin = -\f{U}{2N} \v S\c\v S + J\phi \sum_q \sin\p{q} s_{\z,q},
   \label{eq:H_spin}
@@
 There are three important observations to make about these results.
-First, the fact that $H_{\t{3LD}}$ acts identically on all spins means we can freely replace the site index $j$ with a momentum index $q$ (as can be verified by substituting $c_{j\mu}=\frac1{\sqrt{L}}\sum_k e^{-\i q\c j} c_{q\mu}$), which is important to ensure that this drive addresses the same spin degrees of freedom as the spin Hamiltonian $H_{\t{spin}}$ in Eq.~\eqref{eq:H_spin}.
+First, the fact that $H_{\t{3LD}}$ acts identically on all spins means we can freely replace the site index $j$ with a momentum index $q$ (as can be verified by substituting $c_{j\mu}=\frac1{\sqrt{L}}\sum_k e^{-\i q\c j} c_{q\mu}$), which is important to ensure that this drive addresses the same spin degrees of freedom as the spin Hamiltonian $H_\spin$ in Eq.~\eqref{eq:H_spin}.
 Second, each of $\tilde\Omega_0,\tilde\Omega_+,\tilde\Omega_-$ can be tuned independently by changing the amplitudes of the driving lasers; some particular Hamiltonians for specific values of these amplitudes are shown in Table \ref{tab:drives}.
@@
 Finally, we comment on the preparation of initial states.
-Initial states are nominally prepared in the ``lab frame'', and must be transformed according to the gauge transformation $c_{q\mu}^\dag\to c_{q-\mu\phi,\mu}^\dag$ prior to evolution under the ``gauged frame'' spin Hamiltonian $H_{\t{spin}}$ in Eq.~\eqref{eq:H_spin} or the three-laser drive $H_{\t{3LD}}$ in Eq.~\eqref{eq:drive_all}.
+Initial states are nominally prepared in the ``lab frame'', and must be transformed according to the gauge transformation $c_{q\mu}^\dag\to c_{q-\mu\phi,\mu}^\dag$ prior to evolution under the ``gauged frame'' spin Hamiltonian $H_\spin$ in Eq.~\eqref{eq:H_spin} or the three-laser drive $H_{\t{3LD}}$ in Eq.~\eqref{eq:drive_all}.
 We assume the capability to prepare an initial state in which all spins are maximally polarized along $\v v_+$, i.e.~$\ket{\v v_+}^{\otimes N}=\ket{s}^{\otimes N}$, which is unaffected by the gauge transformation (up to a global phase).
@@
 
-We now return to the spin Hamiltonian $H_{\t{spin}}$ in Eq.~\eqref{eq:H_spin}, which is valid in the weak-SOC limit $s\phi\ll1$, and henceforth work exclusively in the ``gauged frame'' of $H_{\t{spin}}$ and the three-laser drive $H_{\t{3LD}}$ in Eq.~\eqref{eq:drive_all}.
+We now return to the spin Hamiltonian $H_\spin$ in Eq.~\eqref{eq:H_spin}, which is valid in the weak-SOC limit $s\phi\ll1$, and henceforth work exclusively in the ``gauged frame'' of $H_\spin$ and the three-laser drive $H_{\t{3LD}}$ in Eq.~\eqref{eq:drive_all}.
 At the mean-field (MF) level, up to constant energy shifts the undriven spin Hamiltonian becomes
@@
 As expected, the spin model exhibits a mean-field dynamical phase transition between an ordered phase at small $J\phi/U$ and a disordered phase at large $J\phi/U$.
-The ordered phase has a non-zero magnetization $\sigma_\MF$ and an interaction energy $\bbk{\ss}_\MF$ that asymptotically approach their maximal values as $J\phi/U\to0^+$.
+The ordered phase has a non-zero magnetization $\sigma_\MF$ and an interaction energy $\bbk{\ss}_\MF$ that asymptotically approach their maximal values as $J\phi/U\to0$.
 The disordered phase has no (time-averaged) magnetization, $\sigma_\MF=0$, but the interaction energy $\bbk{\ss}_\MF$ nonetheless indicates persistent nontrivial inter-spin correlations when $n>2$.
 These nontrivial correlations vanish as $J\phi/U\to\infty$, in which case $\bbk{\ss}_\MF$ approaches the minimal value allowed by conservation laws (clarified below).
-By minimizing the reduced field $J\phi/U$ for which $\sigma_\MF=0$, we numerically find that the transition between ordered and disordered phases occurs at a critical field $\p{J\phi/U}_{\t{crit}}=\p{n/2}^{-\alpha}$ with $\alpha\approx1/3$ (see Figure \ref{fig:crit_fields_X}).
+By minimizing the reduced field $J\phi/U$ for which $\sigma_\MF=0$, we numerically find that the transition between ordered and disordered phases occurs at a critical field $\p{J\phi/U}_\crit=\p{n/2}^{-\alpha}$ with $\alpha\approx1/3$ (see Figure \ref{fig:crit_fields_X}).
+When $n=2$, this transition is consistent with the predictions of a Lax vector analysis \cite{yuzbashyan2005nonequilibrium, yuzbashyan2006dynamical, yuzbashyan2006relaxation, yuzbashyan2015quantum, smale2019observation} that exploits integrability of $H_\spin$ to determine long-time behavior.
+However, a straightforward application of the same analysis fails to predict the mean-field transition for $n>2$, so additional theoretical tools are necessary to understand this transition.
+We elaborate on this point in Appendix \ref{sec:lax}.
 
@@
 \caption{
-The critical value of $\p{J\phi/U}_{\t{crit}}$ as determined by mean-field simulations of $N=100$ spins initially in the $x$-polarized state $\ket\X$.
-A single-parameter fit to $\p{J\phi/U}_{\t{crit}}=\p{n/2}^{-\alpha}$ finds $\alpha=0.333(5)$, and $\alpha=1/3$ is consistent with all mean-field results to within an uncertainty determined by the resolution of $J\phi/U$ in mean-field simulations.
+The critical value of $\p{J\phi/U}_\crit$ as determined by mean-field simulations of $N=100$ spins initially in the $x$-polarized state $\ket\X$.
+A single-parameter fit to $\p{J\phi/U}_\crit=\p{n/2}^{-\alpha}$ finds $\alpha=0.333(5)$, and $\alpha=1/3$ is consistent with all mean-field results to within an uncertainty determined by the resolution of $J\phi/U$ in mean-field simulations.
 }
@@
 \end{align}
-The rescaling of magnetization and interaction energy can be understood by considering their limiting behavior as $J\phi/U\to\infty$ or $J\phi/U\to0^+$.
+The rescaling of magnetization and interaction energy can be understood by considering their limiting behavior as $J\phi/U\to\infty$ or $J\phi/U\to0$.
 
@@
 
-In the weak-field limit $J\phi/U\to0^+$, the spin-locking $\v S\c\v S$ interactions of the spin Hamiltonian $H_{\t{spin}}$ energetically restrict dynamics to the permutationally-symmetric (PS) manifold.
+In the weak-field limit $J\phi/U\to0$, the spin-locking $\v S\c\v S$ interactions of the spin Hamiltonian $H_\spin$ energetically restrict dynamics to the permutationally symmetric (PS) manifold.
 At second order in perturbation theory, the effective spin Hamiltonian within the PS manifold becomes (see Appendix \ref{sec:pert_theory})
 \begin{align}
-  H_{\t{spin}}^{\t{eff}}
-  = \f{\p{J\phi}^2}{\p{N-1}U}
+  H_\spin^\eff
+  = \f{\p{J\phi}^2}{2\p{N-1}U}
   \times \sp{S_\z^2 - N\sum_q s_{\z,q}^2},
+  \label{eq:H_spin_eff}
 \end{align}
@@
 \begin{align}
-  H_{\t{MFT}}^{\t{eff}} = -\f{\p{J\phi}^2}{U} \sum_q s_{\z,q}^2,
-\end{align}
-where we have used the fact that the axial magnetizations $\bk{s_{\z,q}}=\frac1N\bk{S_\z}$ within the PS manifold, and $\bk{S_\z}=0$ is conserved by $H_{\t{spin}}$.
-The weak-field effective Hamiltonian preserves permutational symmetry, so $\bbk{\ss}_\MF\to1$ as $J\phi/U\to0^+$.
-Moreover, both the spin Hamiltonian $H_{\t{spin}}$ and the initial state $\ket\X$ are invariant (up to global phase) under the action of $\Pi R_\x^\pi$, where $R_\x^\theta \equiv e^{-\i\theta S_\x}$ is a rotation about the $x$ axis by $\theta$ and $\Pi\equiv\sum_{q>0}\v s_q\c\v s_{-q}$ is a permutation of spins $q\leftrightarrow-q$, which is to say that
-\begin{align}
-  e^{-\i t H_{\t{spin}}} \ket\X
-  &\simeq \p{\Pi R_\x^\pi e^{-\i t H_{\t{spin}}} {R_\x^\pi}^\dag \Pi^\dag} \Pi R_\x^\pi \ket\X
+  H_\MF^\eff = -\f{\p{J\phi}^2}{2U} \sum_q s_{\z,q}^2,
+\end{align}
+where we have used the fact that the axial magnetizations $\bk{s_{\z,q}}=\frac1N\bk{S_\z}$ within the PS manifold, and the initial value of $\bk{S_\z}=0$ is conserved by $H_\spin$.
+The weak-field effective Hamiltonian preserves permutational symmetry, so $\bbk{\ss}_\MF\to1$ as $J\phi/U\to0$.
+Moreover, both the spin Hamiltonian $H_\spin$ and the initial state $\ket\X$ are invariant (up to global phase) under the action of $\Pi_\z R_\x^\pi$, where $R_\x^\theta \equiv e^{-\i\theta S_\x}$ is a rotation about the $x$ axis by $\theta$ and $\Pi_\z\equiv\sum_{q>0}\v s_q\c\v s_{-q}$ is a permutation of spins $q\leftrightarrow-q$\footnote{$\Pi_\z$ can also be thought of a reflection operator that inverts momentum along the $z$ axis.}, which is to say that
+\begin{align}
+  e^{-\i t H_\spin} \ket\X
+  &\simeq \p{\Pi_\z R_\x^\pi e^{-\i t H_\spin} {R_\x^\pi}^\dag \Pi_\z^\dag} \Pi_\z R_\x^\pi \ket\X
   \label{eq:sym_x_start} \\
-  &\simeq \Pi R_\x^\pi e^{-\i t H_{\t{spin}}} \ket\X,
+  &\simeq \Pi_\z R_\x^\pi e^{-\i t H_\spin} \ket\X,
 \end{align}
@@
   \bk{S_\y}
-  = \bk{{R_\x^\pi}^\dag \Pi^\dag S_\y \Pi R_\x^\pi}
+  = \bk{{R_\x^\pi}^\dag \Pi_\z^\dag S_\y \Pi_\z R_\x^\pi}
   = -\bk{S_\y} = 0
@@
 \begin{align}
-  \lim_{J\phi/U\to0^+} \sigma_\MF
+  \lim_{J\phi/U\to0} \sigma_\MF
   = \abs{\lim_{T\to\infty} \f1T \int_0^T \dd\tau
@@
 \begin{align}
-  \lim_{J\phi/U\to0^+} \sigma_\MF
+  \lim_{J\phi/U\to0} \sigma_\MF
   = \f1{2\pi} \int_0^{2\pi} \dd\tau \p{\cos\tau}^{n-2}
@@
 \end{align}
+When going beyond mean-field theory, correlations induced by $S_\z^2$ in Eq.~\eqref{eq:H_spin_eff} will cause $\bk{S_\x}$ to decay on time scales that diverge as $N\to\infty$, so the mean-field magnetization $\sigma_\MF$ should be interpreted as the ``intermediate-time-average'' of $\vec\sigma$.
 
@@
 \end{align}
-Similarly to Figure \ref{fig:mean_mag_int_X}, Figures \ref{fig:mean_mag_int_XX} and \ref{fig:mean_mag_int_XXI} respectively show the time-averaged magnetization $\sigma_\MF$ and interaction energy $\bbk{\ss}_\MF$ throughout mean-field dynamics of the initial states $\ket\XX$ and $\ket\XXI$.
-These figures exclude the trivial case of spin dimension $n=2$, for which $\ket\xx=\ket{-s}$ is an eigenstate of $H_{\t{spin}}$ and $\ket\xxi=e^{-\i\frac{\pi}{2}s_\z}\ket\x\equiv\ket\y$.
+Similarly to Figure \ref{fig:mean_mag_int_X}, Figures \ref{fig:mean_mag_int_XX} and \ref{fig:mean_mag_int_XXI} show the time-averaged magnetization $\sigma_\MF$ and interaction energy $\bbk{\ss}_\MF$ throughout mean-field dynamics of the initial states $\ket\XX$ and $\ket\XXI$.
+These figures exclude the trivial case of spin dimension $n=2$, for which $\ket\xx=\ket{-s}$ is an eigenstate of $H_\spin$ and $\ket\xxi=e^{-\i\frac{\pi}{2}s_\z}\ket\x\equiv\ket\y$ is spin-polarized along $y$.
 The first and perhaps most interesting observation to make about Figures \ref{fig:mean_mag_int_XX} and \ref{fig:mean_mag_int_XXI} is that they are different, signifying the importance of intra-spin coherences for the dynamical behavior of multilevel spin models.
 
-Unlike the case of an initial spin-polarized state $\ket\X$, the mean-field dynamics of the initial multi-cat state $\ket\XX$ (Figure \ref{fig:mean_mag_int_XX}) exhibit no sharp transition between distinct dynamical phases: the time-averaged magnetization $\sigma_\MF=0$ for all values of the field $J\phi/U$, and the interaction energy $\bbk{\ss}_\MF$ smoothly crosses over from a maximal value of 1 to a minimal value of $2\gamma\p{n}$.
-The minimal value of $\bbk{\ss}_\MF$ approached as $J\phi/U\to\infty$ can be explained with arguments identical to those above Eq.~\eqref{eq:lim_ss}, which now imply that
+Unlike Figure \ref{fig:mean_mag_int_X} (for $\ket\X$), Figure \ref{fig:mean_mag_int_XX} (for $\ket\XX$) exhibits no sharp transition between distinct dynamical phases: the time-averaged magnetization $\sigma_\MF=0$ for all values of the field $J\phi/U$, and the interaction energy $\bbk{\ss}_\MF$ smoothly crosses over from a maximal value of 1 to a minimal value of $2\gamma\p{n}$.
+The minimal value of $\bbk{\ss}_\MF$ approached as $J\phi/U\to\infty$ can be explained with arguments identical to those above and below Eq.~\eqref{eq:lim_ss}, which now imply that
 \begin{align}
@@
 \end{align}
-The vanishing magnetization $\sigma_\MF=0$ in Figure \ref{fig:mean_mag_int_XX} is protected by symmetries of $H_{\t{spin}}$ and $\ket\XX$.
-For all initial states that we have considered, the value of $\bk{S_\z}=0$ is conserved by the spin Hamiltonian $H_{\t{spin}}$.
-Moreover, both the spin Hamiltonian $H_{\t{spin}}$ and the state $\ket\XX$ are invariant (up to global phase) under the action of $R_\z^\pi$, where $R_\z^\theta\equiv e^{-\i\theta S_\z}$.
+The vanishing magnetization $\sigma_\MF=0$ in Figure \ref{fig:mean_mag_int_XX} is protected by symmetries of $H_\spin$ and $\ket\XX$.
+For all initial states that we have considered, the value of $\bk{S_\z}=0$ is conserved by the spin Hamiltonian $H_\spin$.
+Moreover, both the spin Hamiltonian $H_\spin$ and the state $\ket\XX$ are invariant (up to global phase) under the action of $R_\z^\pi$, where $R_\z^\theta\equiv e^{-\i\theta S_\z}$.
 Similarly to Eqs.~\eqref{eq:sym_x_start}--\eqref{eq:sym_x_end}, this symmetry implies that
@@
 \begin{align}
-  \ket\XXI \simeq R_\z^{\pi/2} T_\z^{\pi/2} \ket\X,
+  \ket\XXI \simeq T_\z^{\pi/2} R_\z^{\pi/2} \ket\X,
 \end{align}
 where $T_\z^\theta\equiv e^{-\i\theta S_\z^2}$.
-The operators $R_\z^\theta$ and $T_\z^\theta$ are generated by axial fields that respect permutational symmetry, and therefore commute with the spin Hamiltonian $H_{\t{spin}}$, so
-\begin{align}
-  e^{-\i t H_{\t{spin}}} \ket\XXI
-  &\simeq e^{-\i t H_{\t{spin}}} R_\z^{\pi/2} T_\z^{\pi/2} \ket\X \\
-  &\simeq R_\z^{\pi/2} T_\z^{\pi/2} e^{-\i t H_{\t{spin}}} \ket\X.
+The operators $R_\z^\theta$ and $T_\z^\theta$ are generated by axial fields that respect permutational symmetry, and therefore commute with the spin Hamiltonian $H_\spin$, so
+\begin{align}
+  e^{-\i t H_\spin} \ket\XXI
+  &\simeq e^{-\i t H_\spin} T_\z^{\pi/2} R_\z^{\pi/2} \ket\X \\
+  &\simeq T_\z^{\pi/2} R_\z^{\pi/2} e^{-\i t H_\spin} \ket\X.
 \end{align}
@@
 \begin{align}
-  {T_\z^\theta}^\dag {R_\z^\theta}^\dag \, \ss \, R_\z^\theta T_\z^\theta
+  {R_\z^\theta}^\dag {T_\z^\theta}^\dag \, \ss \, T_\z^\theta R_\z^\theta
   = \ss,
 \end{align}
-which implies that the interaction energy $\bk{\ss}$ throughout dynamics of the initial multi-cat state $\ket\XXI$ is the same as that of the initial spin-polarized state $\ket\X$.
-
-To make sense of why the magnetization $\sigma_\MF$ should behave identically in Figure \ref{fig:mean_mag_int_XXI} for $\ket\XXI$ as in Figure \ref{fig:mean_mag_int_X} for $\ket\X$, we can ignore the time-independent factors $R_\z^\theta$, which only rotate the mean spin vector $\bbk{\vec\sigma}_\MF$ without changing its magnitude.
-
-\red{[explain $\sigma_\MF$ is identical in Figures \ref{fig:mean_mag_int_X} and \ref{fig:mean_mag_int_XXI}]}
+which implies that the interaction energy $\bk{\ss}$ throughout dynamics of the initial multi-cat state $\ket\XXI$ is the same as that of the spin-polarized state $\ket\X$.
+
+To make sense of why the magnetization $\sigma_\MF$ is identical in Figure \ref{fig:mean_mag_int_XXI} for $\ket\XXI$ as in Figure \ref{fig:mean_mag_int_X} for $\ket\X$, we follow a four-part argument:
+\begin{enumerate}
+\item \label{pt:spin_mat} The time-averaged magnetization vector $\bbk{\vec\sigma}_\MF$ can be written as a function of the time-averaged spin matrix $\bbk{\bar{\v s}}_\MF$.
+\item \label{pt:diags} The spin matrix $\bbk{\bar{\v s}}_\MF$ is only ever nonzero on its diagonal and anti-diagonal, regardless of the initial state.
+That is, nonzero components $\bbk{\bar s_{\mu\nu}}_\MF$ of $\bbk{\bar{\v s}}_\MF$ always have $\mu=\pm\nu$ (see discussion below).
+\item \label{pt:twist} The twist operator $T_\z^\theta$ acts trivially on the diagonal and anti-diagonal components of $\bar{\v s}$, which together with point \ref{pt:diags} implies that $\bbk{{T_\z^\theta}^\dag \, \bar{\v s} \, T_\z^\theta}_\MF = \bbk{\bar{\v s}}_\MF$.
+\item \label{pt:rot} The rotation operator $R_\z^\theta$ merely rotates the magnetization vector $\bbk{\vec\sigma}_\MF$ without changing its magnitude.
+\end{enumerate}
+Altogether, points \ref{pt:spin_mat}--\ref{pt:rot} imply that the magnetization
+\begin{align}
+  \sigma_\MF = \abs{\bbk{\vec\sigma}_\MF}
+  = \abs{\bbk{{R_\z^\theta}^\dag {T_\z^\theta}^\dag \,
+      \vec\sigma \, T_\z^\theta R_\z^\theta}_\MF}
+\end{align}
+is the same for the initial state $\ket\XXI$ as for $\ket\X$.
+
+The only nontrivial step in the above argument is point \ref{pt:diags}, which says that $\bbk{\bar s_{\mu\nu}}_\MF$ is guaranteed to be zero unless $\mu=\pm\nu$.
+This observation, nominally a numerical result of mean-field simulations, can be understood as follows.
+The eigenstates $\ket{m,w}$ of $H_\spin$ are uniquely identified by definite numbers $m=(m_s,m_{s-1},\cdots,m_{-s})$ of atoms occupying each internal spin state $\mu\in\set{s,s-1,\cdots,-s}$, and an auxiliary index $w$ that encodes how $\ket{m,w}$ transforms under permutations of all spins (see Appendix \ref{sec:pert_theory})\footnote{Seen otherwise, since $S_{\mu\mu}$ commutes with $H_\spin$, eigenvectors of $H_\spin$ can be indexed by eigenvalues of $S_{\mu\mu}$.
+The number $m_\mu$ is then the eigenvalue of $\ket{m,w}$ with respect to $S_{\mu\mu}$, i.e.~$S_{\mu\mu}\ket{m,w} = m_\mu\ket{m,w}$, while $w$ encodes all other information required to uniquely specify $\ket{m,w}$.}.
+The operator $\bar s_{\mu\nu}=\frac1N S_{\mu\nu}$ with $\mu\ne\nu$ couples the state $\ket{m,w}$ to states $\ket{m',w'}$ in which $(m_\mu',m'_\nu)=(m_\mu+1,m_\nu-1)$.
+Generically, states $\ket{m,w}$ and $\ket{m',w'}$ with $m\ne m'$ will have different energies, so their coherence averages to zero when evaluating time-averaged expectation values.
+
+However, degeneracies yield stationary (time-independent) coherences that survive time-averaging.
+In the weak-field limit $J\phi/U\to0$, such a degeneracy occurs at the mean-field level between PS states differing only in the populations $m_\mu,m_{-\mu}$ (with a fixed value of $m_\mu+m_{-\mu}$), as the effective Hamiltonian becomes $H_\MF^\eff\propto\sum_\mu\mu^2m_\mu$.
+This symmetry is preserved at all orders in perturbation theory\footnote{Only even powers of the ``perturbation'' $\sum_q\sin\p{q}s_{\z,q}$ can be nonzero within the PS manifold, and even powers of this perturbation exhibit the same mean-field degeneracy between states differing only in the populations $m_\mu,m_{-\mu}$.}, so some coherence between such states is preserved as $(J\phi/U)\to(J\phi/U)_\crit$, although this coherence decays as perturbative corrections to degenerate eigenstates cause them to leak out of the PS manifold (and thereby have a smaller overlap with the initial state $\ket\X$).
+Note that beyond-mean-field effects break the symmetry protecting anti-diagonal components of $\bbk{\v s}_\MF$, causing them to decay on time scales that should diverge as $N\to\infty$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusions}
+
+\red{[to be completed]}
 
@@
 
-\red{[to be completed]}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Restricting spin operators to the permutationally-symmetric manifold}
+Here we work out a general perturbation theory for SU($n$) ferromagnets with a gapped permutationally symmetric (PS) manifold.
+We begin with an SU($n$)-symmetric interaction Hamiltonian of the form
+\begin{align}
+  H_0 = \sum_{i<j} g_{ij} \, \Pi_{ij},
+  &&
+  \Pi_{ij} \equiv \v s_i\c\v s_j
+  = \sum_{\mu,\nu} s_{\mu\nu i} s_{\nu\mu j},
+  \label{eq:H_0}
+\end{align}
+where $g_{ij}$ are (real) scalar coefficients for the permutation operators $\Pi_{ij}$, and $s_{\mu\nu i}\equiv c_{\mu i}^\dag c_{\nu i}$ is a transition operator for spin $i$.
+We can then consider the addition of, for example, an inhomogeneous magnetic field or Ising couplings,
+\begin{align}
+  H_{\t{field}} = \sum_i B_i s_{\z,i},
+  &&
+  H_{\t{Ising}} = \sum_{i\ne j} J_{ij} s_{\z,i} s_{\z,j},
+  \label{eq:field_ising}
+\end{align}
+or more generally an $M$-body operator
+\begin{align}
+  \O\p{w,X} = \sum_{k\in\D_N\p{M}} w_k X_k,
+\end{align}
+where $w$ is a dimension-$M$ (i.e.~$M$-index) tensor of scalar
+coefficients $w_k\equiv w_{k_1k_2\cdots k_M}$; $X$ is an $M$-spin
+operator, e.g.~$s_\z\otimes s_\z$ in the case of Ising interactions
+with $M=2$; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is a list of the
+individual spins $k_i\in\ZZ_N\equiv\set{1,2,\cdots,N}$ that the operator $X_k\equiv X_{k_1k_2\cdots k_M}$ acts on; and
+\begin{align}
+  \D_N\p{M} \equiv
+  \set{ k \in \ZZ_N^M : \t{all entries $k_i$ of $k$ are distinct} },
+\end{align}
+is the strictly ``off-diagonal'' part of $\ZZ_N^M$, which is necessary to identify for a consistent definition of $X_k$ as an $M$-body operator.
+In this notation, the magnetic field and Ising Hamiltonians in Eq.~\eqref{eq:field_ising} respectively become $\O\p{B,s_{\z}}$ and $\O\p{J,s_\z\otimes s_\z}$.
+
+If the addition $\O\p{w,X}$ to the SU($n$)-symmetric Hamiltonian $H_0$ in Eq.~\eqref{eq:H_0} is sufficiently small, namely with operator norm $\norm{\O\p{w,X}}$ less than half the spectral gap $\Delta_{\t{gap}}$ of $H_0$, $\norm{\O\p{w,X}}<\Delta_{\t{gap}}/2$, then we can treat the effect of $\O\p{w,X}$ on the ground-state PS manifold $\E_0$ perturbatively.
+The effective Hamiltonians $H_\eff^{(1)}$ and $H_\eff^{(2)}$ induced by $\O\p{w,X}$ on the PS manifold $\E_0$ at leading orders in perturbation theory are \cite{bravyi2011schrieffer}
+\begin{align}
+  H_\eff^{(1)} = \P_0 \O\p{w,X} \P_0,
+  &&
+  H_\eff^{(2)} = - \sum_{\Delta\ne0}
+  \f1\Delta \P_0 \O\p{w,X} \P_\Delta \O\p{w,X} \P_0,
+  \label{eq:H_eff}
+\end{align}
+where $\P_\Delta$ is a projector onto the eigenspace $\E_\Delta$ of $H_0$ with interaction energy $\Delta$ above that of the PS manifold.
+The first order effective Hamiltonian $H_\eff^{(1)}$ simply projects $\O\p{w,X}$ onto the PS manifold $\E_0$, and takes the form
+\begin{align}
+  H_\eff^{(1)} = \mean w \, \col X,
+  \label{eq:H_eff_1}
+\end{align}
+where the coefficient $\mean w$ is the average of all coefficients $w_k$; and $\col X$ is a collective version of $X$:
+\begin{align}
+  \mean w \equiv \f1{\abs{\D_N\p{M}}}
+  \sum_{k\in\D_N\p{M}} w_k,
+  &&
+  \col X \equiv \sum_{k\in\D_N\p{M}} X_k,
+  \label{eq:mean_col}
+\end{align}
+with $\abs{\D_N\p{M}}=\prod_{j=0}^{M-1}\p{N-j}$.
+In the case of a magnetic field $s_\z$ or Ising interactions  $s_\z\otimes s_\z$, for example,
+\begin{align}
+  \col{s_\z} = \sum_i s_\z^{(i)} = S_\z,
+  &&
+  \col{s_\z\otimes s_\z}
+  = \sum_{i\ne j} s_\z^{(i)} s_\z^{(j)}
+  = S_\z^2 - N \sum_i s_{\z,i}^2.
+\end{align}
+The second order effective Hamiltonian $H_\eff^{(2)}$ in Eq.~\eqref{eq:H_eff} takes more work to simplify due to the presence of a projector $\P_\Delta$ onto the manifold $\E_\Delta$ of states with excitation energy $\Delta$.
+This projector essentially picks off the part of $\O\p{w,X}$ that is strictly off-diagonal with respect to the ground- and excited-state manifolds $\E_0$ and $\E_\Delta$.
+We therefore need to decompose $\O\p{w,X}$ into components that generate states of definite excitation energy when acting on PS states $\ket\psi\in\E_0$.
+The SU($n$) symmetry of $H_0$ enables such a decomposition to take the form
+\begin{align}
+  H_0 \O\p{w,X} \ket\psi
+  = \sum_\Delta \p{E_0+\Delta} \O\p{w^\Delta,X} \ket\psi,
+  &&
+  E_0 \equiv \sum_{i<j} g_{ij},
+  \label{eq:excitation}
+\end{align}
+where $E_0$ is the interaction energy of PS states, and thinking of the tensor $w$ as a $\abs{\D_N\p{M}}$-component vector, the tensor $w^\Delta$ can be found by
+\begin{enumerate*}
+\item using the coefficients $g_{ij}$ to construct a matrix $g^{(M)}$ of dimensions $\abs{\D_N\p{M}}\times\abs{\D_N\p{M}}\sim N^M\times N^M$, and
+\item projecting $w$ onto the eigenspace of $g^{(M)}$ with eigenvalue $\Delta$.
+\end{enumerate*}
+We construct $g^{(M)}$ for the single-body ($M=1$) case in Appendix \ref{sec:eigenstates}, and provide explicit forms of $g^{(M)}$ with arbitrary $M$.
+
+Equipped with the decomposition $\O\p{w,X}=\sum_\Delta\O\p{w^\Delta,X}$ with terms $\O\p{w^\Delta,X}$ that generate states of definite excitation energy $\Delta$, we can expand
+\begin{align}
+  H_{\t{eff}}^{(2)} = -\sum_{\Delta\ne0} \f1\Delta
+  \P_0 \O\p{w^\Delta,X}^2 \P_0.
+\end{align}
+If $X$ is a single-body operator, then
+\begin{align}
+  H_{\t{eff}}^{(2)}
+  = \sum_{\Delta\ne0} \f{w^\Delta\c w^\Delta}{N\p{N-1}\Delta}
+  \p{\col{X}^2 - N \col{X^2}},
+\end{align}
+and if furthermore all $g_{ij}=-U/N$, as for $H_{\t{int}}$ in Eq.~\eqref{eq:H_int}, then the only relevant excitation energy is $\Delta=U$, and
+\begin{align}
+  w^U\c w^U = \sum_i \p{w_i-\mean w}^2 = N \widetilde{w}^2
+\end{align}
+is simply $N$ times the variance $\widetilde{w}^2$ of $w$, so
+\begin{align}
+  H_{\t{eff}}^{(2)}
+  = \f{\widetilde{w}^2}{\p{N-1}U} \p{\col{X}^2 - N \col{X^2}}.
+\end{align}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Generating excitation energy eigenstates}
+\label{sec:eigenstates}
+
+Here we construct the matrix $g^{(M)}$ that enables decomposing $M$-body operators $\O\p{w,X}$ into terms $\O\p{w^\Delta,X}$ that generate states of definite excitation energy $\Delta$ above the PS manifold, as in Eq.~\eqref{eq:excitation}.
+We work through the calculation of $g^{(1)}$ explicitly, and provide the result for $g^{(M)}$ from a generalized version of the same calculation.
+To this end, we consider the action of a single-body operator $\O\p{w,X}=\sum_iw_iX_i$ on an arbitrary PS state $\ket\psi\in\E_0$ and expand
+\begin{align}
+  H_0 \O\p{w,X} \ket\psi
+  = \f12 \sum_{i\ne j} \sum_k g_{ij} w_k \Pi_{ij} X_k \ket\psi,
+  \label{eq:single_body_diagnosis_start}
+\end{align}
+where strictly speaking $g_{ij}$ has only been defined for $i<j$, so for completeness we define $g_{ji}=g_{ij}$ and $g_{ii}=0$.
+The sum in Eq.~\eqref{eq:single_body_diagnosis_start} has terms with $k\in\set{i,j}$ and terms with $k\notin\set{i,j}$.
+In the case of $k\notin\set{i,j}$, the permutation operator $\Pi_{ij}$ commutes with $X_k$ and annihilates on $\ket\psi$, and we can replace the sum
+\begin{align}
+  \sum_{k\notin\set{i,j}} \to \sum_k - \sum_{k\in\set{i,j}},
+\end{align}
+allowing us to simplify
+\begin{align}
+  \f12 \sum_{i\ne j} \sum_{k\notin\set{i,j}}
+  g_{ij} w_k \Pi_{ij} X_k \ket\psi
+  = E_0 \O\p{w,X} \ket\psi
+  - \f12 \sum_{i\ne j} \sum_{k\in\set{i,j}} g_{ij} w_k X_k \ket\psi,
+\end{align}
+where $E_0=\frac12\sum_{i\ne j}g_{ij}$ is the interaction energy the PS state $\ket\psi\in\E_0$.
+Switching the order of sums over $i\ne j$ and $k\in\set{i,j}$ as
+\begin{align}
+  \sum_{i\ne j} \sum_{k\in\set{i,j}}
+  \to \sum_k \sum_{\substack{i\ne j\\\set{i,j}\ni k}},
+\end{align}
+we can simplify
+\begin{align}
+  \f12 \sum_{\substack{i\ne j\\\set{i,j}\ni k}} g_{ij}
+  = \f12 \sum_i g_{ik} + \f12 \sum_j g_{kj}
+  = g_k,
+  &&
+  g_k \equiv \sum_i g_{ik},
+\end{align}
+which implies that the terms in Eq.~\eqref{eq:single_body_diagnosis_start}
+with $k\notin\set{i,j}$ are
+\begin{align}
+  \f12 \sum_{i\ne j} \sum_{k\notin\set{i,j}}
+  g_{ij} w_k \Pi_{ij} X_k \ket\psi
+  = E_0 \O\p{w,X} \ket\psi - \sum_k g_k w_k X_k \ket\psi.
+\end{align}
+The terms in Eq.~\eqref{eq:single_body_diagnosis_start} with
+$k\in\set{i,j}$, meanwhile, are
+\begin{align}
+  \f12 \sum_{\substack{i\ne j\\k\in\set{i,j}}}
+  g_{ij} w_k \Pi_{ij} X_k \ket\psi
+  = \sum_{i,j} g_{ij} w_j X_i \ket\psi.
+\end{align}
+so in total
+\begin{align}
+  H_0 \O\p{w,X} \ket\psi
+  = E_0 \O\p{w,X} \ket\psi
+  + \sum_k \sp{\sum_j g_{kj} w_j - g_k w_k} X_k \ket\psi.
+\end{align}
+The action of the single-body perturbation $\O\p{w,X}$ on a permutationally symmetric state therefore generates an eigenstate of $H_0$ with interaction energy $E_0+\Delta$ if the vector $w=\sum_k w_k \ket{k}$ satisfies the eigenvalue equation
+\begin{align}
+  g^{(1)} \c w = \Delta w, && g^{(1)} \equiv g - \diag\vec g,
+  \label{eq:single_body_eig}
+\end{align}
+where $g \equiv\sum_{i,j} g_{ij} \op{i}{j}$ is a matrix of all couplings $g_{ij}$; the vector $\vec g \equiv \sum_{i,j} g_{ij} \ket{i} = \sum_i g_i \ket{i}$ is the sum of all columns of $g$; and the matrix $\diag\vec g \equiv \sum_i g_i \op{i}$ has $\vec g$ on the diagonal and zeroes everywhere else.
+
+A similar calculation as above with arbitrary $M$ yields an eigenvalue equation of the form
+\begin{align}
+  g^{(M)} \c w = \Delta w,
+  \label{eq:multi_body_problem}
+\end{align}
+where we treat $w$ as an $\abs{\D_N\p{M}}$-component vector, and $g^{(M)}$ is a matrix with dimensions $\abs{\D_N\p{M}}\times\abs{\D_N\p{M}}$.
+In the case of $M=2$, we have
+\begin{align}
+  g^{(2)} = \sum_{\p{k,\ell}\in\D_N\p{2}} \ket{k\ell}
+  \sp{\sum_{\substack{i\in\ZZ_N\\i\notin\set{k,\ell}}}
+    \p{ g_{ik} \bra{i\ell} + g_{i\ell} \bra{ki} }
+    + g_{k\ell} \bra{\ell k}
+    - \p{g_k + g_\ell - g_{k\ell}} \bra{k\ell}},
+  \label{eq:two_body_mat}
+\end{align}
+and more generally
+\begin{align}
+  g^{(M)} = \sum_{k\in\D_N\p{M}} \ket{k}
+  \sp{\sum_{a\in\ZZ_M} \sum_{\substack{i\in\ZZ_N\\i\notin k}}
+    g_{ik_a} \bra{k_{a:i}}
+    + \sum_{\set{a,b}\in\C_M\p{2}} g_{k_ak_b} \bra{k_{a\leftrightarrow b}}
+    - \tilde g_k \bra{k}},
+  \label{eq:multi_body_mat}
+\end{align}
+where $k_a\in k=\p{k_1,k_2,\cdots,k_M}$; $k_{a:i}$ a list that is equal to $k$ except at the $a$-th position, where $k_a$ replaced is by $i$, i.e.~$k_{a:i}=\p{\cdots,k_{a-1},i,k_{a+1},\cdots}$; $\C_L\p{p}$ is the set of all subsets (``choices'') of $p$ elements from $\ZZ_L$; $k_{a\leftrightarrow b}$ is equal to $k$ except at the $a$-th and $b$-th positions, at which $k_a$ and $k_b$ are switched; and
+\begin{align}
+  \tilde g_k
+  \equiv \sum_{\substack{\set{i,j}\in\C_N\p{2}\\i\in k~\t{or}~j\in k}} g_{ij}
+  = \sum_{i\in k} g_i - \sum_{\set{a,b}\in\C_M\p{2}} g_{k_ak_b}.
+\end{align}
+If the tensor $w$ is permutationally symmetric, meaning that $w_k$ is invariant under arbitrary permutations of $k$, then this symmetry is preserved by $g^{(M)}$.
+In this case, we can replace sums over $k\in\D_N\p{M}$ in Eqs.~\eqref{eq:two_body_mat} and \eqref{eq:multi_body_mat} by sums over $k\in\C_N\p{M}$, and replace vectors $\ket{k_1,k_2,\cdots,k_M}\to\ket{\set{k_1,k_2,\cdots,k_M}}$, such that e.g.~$\ket{k_{a\leftrightarrow b}}=\ket{k}$.
+These replacements reduce the size of $g^{(M)}$ from $\abs{\D_N\p{M}}\times\abs{\D_N\p{M}}$ to $\abs{\C_N\p{M}}\times\abs{\C_N\p{M}}$, where $\abs{\D_N\p{M}}=\prod_{j=0}^{M-1}\p{N-j}=M!\times{N\choose M}$ and $\abs{\C_N\p{M}}={N\choose M}$.
+Additional symmetries of $g$ and $w$, such as translational invariance or lattice symmetries, can be used to further reduce the computational complexity of the eigenvalue problem in Eq.~\eqref{eq:multi_body_problem}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Recovering spin-wave theory}
+
+If the interaction Hamiltonian $H_0$ is translationally invariant, then the single-body eigenvalue problem in Eq.~\eqref{eq:single_body_eig} is solvable analytically.
+In this case, the couplings $g_{ij}$ depend only on the separation $\abs{i-j}$, so eigenvectors of $g$ are plane waves of the form
+\begin{align}
+  w_k \equiv \sum_{d\in\ZZ_L^D} e^{\i d\c k} \ket{d},
+\end{align}
+where on a $D$-dimensional periodic lattice of $N=L^D$ spins, lattice sites are indexed by vectors $d\in\ZZ_L^D$, and wavenumbers take on values $k\in\ZZ_L^D\times2\pi/L$.
+The eigenvalues of $g$ can be determined by expanding
+\begin{align}
+  g \c w_k
+  = \sum_{c,d\in\ZZ_L^D} g_{cd} e^{\i d\c k} \ket{c}
+  = \sum_{c,d\in\ZZ_L^D} g_{c,c+d} e^{\i\p{c+d}\c k} \ket{c}
+  = \sum_{d\in\ZZ_L^D} g_{0,d} \cos\p{d\c k} w_k,
+\end{align}
+where the imaginary contributions vanish in the sum over $d$ because $g_{0,d}=g_{0,-d}$.
+The remainder of Eq.~\eqref{eq:single_body_eig} that we need to sort out is $\diag\vec g$, where all $g_i = \sum_{i,j}g_{ij} = \sum_d g_{0,d}$ are equal, which implies that $\diag\vec g = \sum_d g_{0,d}$ is a scalar.
+We thus find that
+\begin{align}
+  g^{(1)} \c w_k = \Delta_k w_k,
+  &&
+  \Delta_k \equiv \sum_{d\in\ZZ_L^D} g_{0,d} \sp{\cos\p{d\c k}-1},
+\end{align}
+in agreement with standard spin-wave theory.
+Excitations generated by the action of $\O\p{w_k,X}$ on PS states $\ket\psi\in\E_0$ are known as spin-waves.
+If $g_{ij}=-U/N$ is constant, then the spin-wave excitation energies are $\Delta_k=U$ independent of the wavenumber $k$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Restricting spin operators to the permutationally symmetric manifold}
 \label{sec:PS_ops}
 
-Here we provide the restriction of a general $M$-body spin operator $\O$ to the permutationally-symmetric (PS) manifold of $N$ spins (each with $n$ internal states).
-Denoting the projector onto the PS manifold by $\P$, our task is essentially to find the coefficients of the expansion
-\begin{align}
-  \P \O_M \P = \sum_{a,b\in\A_n\p{N}} \bk{a|\O_M|b} \op{a}{b},
+Here we provide the restriction of a general $M$-body spin operator $\O$ to the permutationally symmetric (PS) manifold of $N$ spins (each with $n$ internal states).
+Denoting the projector onto the PS manifold by $\P_0$, our task is essentially to find the coefficients of the expansion
+\begin{align}
+  \P_0 \O_M \P_0 = \sum_{a,b\in\A_n\p{N}} \bk{a|\O_M|b} \op{a}{b},
 \end{align}
@@
 We sum over both $\alpha$ and $\beta$ above merely to keep the expression symmetric with respect to transposition $\p{a,\alpha}\leftrightarrow\p{b,\beta}$; in practice, one can simply sum over $\alpha\in\A_n\p{M}$ and set $\beta=b-a+\alpha$, throwing out terms with any $\beta_\mu<0$.
-Note that, by slight abuse of notation, the operator $\O_M$ on the left of \eqref{eq:multi_body_eval} acts on an arbitrary choice of $M$ spins (out of $N$), whereas the operator $\O_M$ on the right of \eqref{eq:multi_body_eval} is simply an $M$-spin operator, with matrix elements $\bk{\alpha|\O_M|\beta}$ evaluated with respect to the PS $M$-spin states $\ket\alpha,\ket\beta\in\A_n\p{M}$.
+Note that, by slight abuse of notation, the operator $\O_M$ on the left of Eq.~\eqref{eq:multi_body_eval} acts on an arbitrary choice of $M$ spins (out of $N$), whereas the operator $\O_M$ on the right of Eq.~\eqref{eq:multi_body_eval} is simply an $M$-spin operator, with matrix elements $\bk{\alpha|\O_M|\beta}$ evaluated with respect to the PS $M$-spin states $\ket\alpha,\ket\beta\in\A_n\p{M}$.
 
@@
 \begin{align}
-  H_{\t{spin}} = -\f{U}{2N}\v S\c\v S + J\phi \sum_q \sin\p{q} s_{\z,q}
+  H_\spin = -\f{U}{2N}\v S\c\v S + J\phi \sum_q \sin\p{q} s_{\z,q}
 \end{align}
@@
 \begin{align}
-  H_{\t{spin}} \to H_{\t{boson}}
+  H_\spin \to H_{\t{boson}}
   = -\f{U}{2N} \sum_{p,q,\mu,\nu}
@@
 \section{Lax vector analysis}
+\label{sec:lax}
 
@@
 \begin{align}
-  H_{\t{spin}}
+  H_\spin
   = -\f{U}{2N} \sum_{\mu,\nu} S_{\mu\nu} S_{\nu\mu}
@@
 where $S_{\mu\nu} = \sum_q s_{\mu\nu q}$.
-The single-body operators involved have normalizations
+The single-body operators that appear in this Hamiltonian have squared norms
 \begin{align}
@@
 \end{align}
-The Lax formulation requires all single-body operators involved to have the same normalization, so we substitute $s_{\tilde\z,q}\equiv s_{\z,q}/\eta$ to expand
-\begin{align}
-  U^{-1} H_{\t{spin}}
+The Lax formulation (following Refs.~\cite{yuzbashyan2005nonequilibrium, yuzbashyan2006dynamical, yuzbashyan2006relaxation, yuzbashyan2015quantum, smale2019observation}) requires all single-body operators involved to have the same normalization, so we substitute $s_{\tilde\z,q}\equiv s_{\z,q}/\eta$ to expand
+\begin{align}
+  \f{H_\spin}{U}
   = -\f1{2N} \sum_{\mu,\nu} S_{\mu\nu} S_{\nu\mu}
@@
 \end{align}
-The normalized Lax vector $\vec\ell\p{z}$ associated with $H_{\t{spin}}$ has components
+The normalized, dimensionless, $(n^2-1)$-component Lax vector $\vec\ell\p{z}$ associated with $H_\spin$, which is defined with an auxiliary complex parameter $z$, has components
 \begin{align}
@@
 \end{align}
-where $\alpha$ indexes elements of basis $\set{s_\alpha}$ of self-adjoint generators of SU($n$), with normalization $\tr\p{s_\alpha^2}=1$.
+where $\alpha$ indexes elements of a basis $\set{s_\alpha}$ of self-adjoint generators of SU($n$), with normalization $\tr\p{s_\alpha^2}=1$.
+The squared magnitude $\vec\ell\p{z}^2=\sum_\alpha\ell_\alpha\p{z}^2$ is a constant of motion (for any $z$), and its residues provide $N$ mutually commuting quantities whose weighted sum recovers $H_\spin$.
+When $n=2$, conservation of these residues provides sufficient dynamical constraints to make the spin system fully integrable.
+In this case, dynamical behavior is governed by the roots of $\vec\ell\p{z}^2$, and the presence (or absence) of complex roots marks distinct dynamical phases of $H_\spin$.
+However, the size of Hilbert space grows with $n$, while the number of conserved quantities provided by the Lax analysis (namely, $N$) does not.
+When $n>2$, there is therefore no guarantee that the roots of $\vec\ell\p{z}^2$ will similarly govern dynamical behavior.
+In fact, a straightforward generalization of the Lax analysis to $n>2$ makes predictions that are inconsistent with the mean-field results in Figures \ref{fig:mean_mag_int_X}--\ref{fig:mean_mag_int_XXI} of the main text.
+We substantiate this claim with a direct calculation of the roots of $\vec\ell\p{z}^2$ below.
+
 Within the permutationally symmetric manifold, we can replace $s_{\alpha,q}\to\bar s_\alpha\equiv\frac1N\sum_q s_{\alpha,q}$ at the cost of $O(1/N)$ errors that vanish as $N\to\infty$, so taking this limit we find
@@
 \begin{align}
-  \I\p{z} \equiv \f1N \sum_q \f1{z-\sin\p{q}}
+  \I\p{z} \equiv \lim_{N\to\infty} \f1N \sum_q \f1{z-\sin\p{q}}
   = \f1{2\pi} \int_0^{2\pi} \f{\dd q}{z-\sin\p{q}}
@@
 \end{align}
-The ``squared norm'' of the Lax vector is therefore
+The squared magnitude of the Lax vector is therefore
 \begin{align}
@@
 \end{align}
-These roots change character when $z=0$, suggesting that the critical field $h_{\t{crit}}$ separating dynamical phases satisfies
-\begin{align}
-  h_{\t{crit}}^2 = \f{Q^2}{\eta^2}.
+These roots change character when $z=0$, suggesting that the critical field $h_\crit$ separating dynamical phases satisfies
+\begin{align}
+  h_\crit^2 = \f{Q^2}{\eta^2}.
 \end{align}
@@
 \begin{align}
-  h_{\t{crit}}^2 = \f{n-1}{n} \times \f{12}{(n+1)n(n-1)}
+  h_\crit^2 = \f{n-1}{n} \times \f{12}{(n+1)n(n-1)}
   = \f{12}{n^2\p{n+1}}.
 \end{align}
-This Lax analysis correctly predicts that $h_{\t{crit}}=1$ when $n=2$, but otherwise predicts $h_{\t{crit}}\sim n^{-3/2}$, which is inconsistent with the finding that $h_{\t{crit}}\sim n^{-1/3}$ in mean-field simulations (see Figure \ref{fig:crit_fields_X} of the main text).
+This Lax analysis correctly predicts that $h_\crit=1$ when $n=2$, but otherwise predicts $h_\crit\sim n^{-3/2}$, which is inconsistent with the finding that $h_\crit\sim n^{-1/3}$ in the mean-field results of the main text (see Figure \ref{fig:crit_fields_X}).
 

### 2021-04-15 sun_phases/spin_model_paper/main.tex
added appendix on three-laser drive with complex amplitudes other minor changes to text

@@
 External fields or additional interactions that respect permutational symmetry can induce nontrivial dynamics within the PS manifold.
-Even if additional Hamiltonian terms break permutational symmetry, as long as the coupling to states outside of the PS manifold is small (compared to the energy gap $U$) these terms can be restricted to the PS manifold perturbatively (see Appendix \ref{sec:pert_theory}) \cite{bravyi2011schrieffer}.
+Even if additional Hamiltonian terms break permutational symmetry, as long as the coupling to states outside of the PS manifold is small compared to the interaction energy $U$, these terms can be restricted to the PS manifold perturbatively (see Appendix \ref{sec:pert_theory}) \cite{bravyi2011schrieffer}.
 Simulating dynamics within the PS manifold requires calculating matrix elements $\bk{\ell|\O|m}$ of spin operators $\O$ with respect to PS states $\ket\ell,\ket{m}$; we discuss this calculation in Appendix \ref{sec:PS_ops}.
@@
 as visualized in Figure \ref{fig:soc_panels}.
-To better interpret this Hamiltonian, we can write it in the simplified form
+To better interpret this Hamiltonian, we can write it in the form
 \begin{align}
@@
 \end{align}
-For two-level spins, $w_{+,q}^{(\phi)}$ is proportional to the identity operator, and $w_{+,q}^{(\phi)}=2\sin\p{\phi/2} s_{q,\z}$, where $s_{\z,q}=\sum_\mu \mu\, s_{\mu\mu q}$ is a spin-$z$ operator for spin $q$, so the kinetic Hamiltonian in the gauged frame describes a (synthetic) inhomogeneous magnetic field:
+For two-level spins, $w_{+,q}^{(\phi)}$ is proportional to the identity operator, and $w_{+,q}^{(\phi)}=2\sin\p{\phi/2} s_{\z,q}$, where $s_{\z,q}=\sum_\mu \mu\, s_{\mu\mu q}$ is a spin-$z$ operator for spin $q$, so the kinetic Hamiltonian in the gauged frame describes a (synthetic) inhomogeneous magnetic field:
 \begin{align}
@@
 In the far-detuned limit $\abs{\Delta}\gg\abs{\Omega_m}$, a second-order perturbative treatment of electronic excitations ($\ket\e$) yields an effective drive Hamiltonian that only addresses ground-state nuclear spins.
-After a gauge transformation $c_{j\mu}^\dag\to e^{\i\phi\mu j} c_{j\mu}^\dag$ (equivalently $s_{mj}\to e^{\i m\phi j}s_{mj}$), we can then make the simplifying assumption that all $\Omega_m$ are real to write
+After additionally making the gauge transformation $c_{j\mu}^\dag\to e^{\i\phi\mu j} c_{j\mu}^\dag$ (equivalently $s_{mj}\to e^{\i m\phi j}s_{mj}$), the drive Hamiltonian then becomes
 \begin{align}
@@
   &&
-  \tilde\Omega_\pm \equiv \f{\Omega_+\pm\Omega_-}{\sqrt\Delta}.
-\end{align}
+  \tilde\Omega_\pm \equiv \f{\Omega_+\pm\Omega_-}{\sqrt\Delta},
+\end{align}
+where we have made the simplifying assumption that all drive amplitudes are real to arrive at the form of $H_{\t{3LD}}^{\t{single}}$ in Eq.~\eqref{eq:drive_single}.
+We relax the assumption of real drive amplitudes in Appendix \ref{sec:full_drive}.
 
@@
 \end{align}
-where $S_\alpha \equiv \sum_q s_{\alpha,q}$, and the time-averaged interaction energy
+where $S_\alpha \equiv \sum_q s_{\alpha,q}$, and the time-averaged (dimensionless) interaction energy
 \begin{align}
@@
 \end{align}
-By design, these non-negative quantities are normalized to a maximal value of one, independent of the system size $N$ or spin dimension $n$.
+By design, these non-negative quantities are normalized to lie on the interval $[0,1]$, independent of the system size $N$ or spin dimension $n$.
 In the remainder of this section we will assume that $n$ is even, both for the sake of experimental relevance (most relevant atomic nuclei are fermionic) and to avoid complications from parity effects.
@@
 \end{align}
-When going beyond mean-field theory, correlations induced by $S_\z^2$ in Eq.~\eqref{eq:H_spin_eff} will cause $\bk{S_\x}$ to decay on time scales that diverge as $N\to\infty$, so the mean-field magnetization $\sigma_\MF$ should be interpreted as the ``intermediate-time-average'' of $\vec\sigma$.
+When going beyond mean-field theory, inter-spin correlations generated by $S_\z^2$ in Eq.~\eqref{eq:H_spin_eff} will cause $\bk{S_\x}$ to decay on time scales that diverge as $N\to\infty$, so the mean-field magnetization $\sigma_\MF$ should be interpreted as an ``intermediate-time-average'' of $\vec\sigma$.
 
@@
 Unlike Figure \ref{fig:mean_mag_int_X} (for $\ket\X$), Figure \ref{fig:mean_mag_int_XX} (for $\ket\XX$) exhibits no sharp transition between distinct dynamical phases: the time-averaged magnetization $\sigma_\MF=0$ for all values of the field $J\phi/U$, and the interaction energy $\bbk{\ss}_\MF$ smoothly crosses over from a maximal value of 1 to a minimal value of $2\gamma\p{n}$.
-The minimal value of $\bbk{\ss}_\MF$ approached as $J\phi/U\to\infty$ can be explained with arguments identical to those above and below Eq.~\eqref{eq:lim_ss}, which now imply that
+The minimal value of $\bbk{\ss}_\MF$ approached as $J\phi/U\to\infty$ can be explained with arguments identical to those in the paragraph containing Eq.~\eqref{eq:lim_ss}, which now imply that
 \begin{align}
@@
 \end{align}
-or more generally an $M$-body operator
+or more generally an $M$-body operator\footnote{At face value, an $M$-body operator with $M>2$ does not typically appear in nature.
+Nonetheless, considering $M>2$ illuminates the structure of eigenstates (and eigenvalues) of $H_0$, and allows us to go to high orders in perturbation theory with single- and two-body perturbations.}
 \begin{align}
@@
 Note that, by slight abuse of notation, the operator $\O_M$ on the left of Eq.~\eqref{eq:multi_body_eval} acts on an arbitrary choice of $M$ spins (out of $N$), whereas the operator $\O_M$ on the right of Eq.~\eqref{eq:multi_body_eval} is simply an $M$-spin operator, with matrix elements $\bk{\alpha|\O_M|\beta}$ evaluated with respect to the PS $M$-spin states $\ket\alpha,\ket\beta\in\A_n\p{M}$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Three-laser drive with complex amplitudes}
+\label{sec:full_drive}
+
+In order to arrive at the single-spin drive Hamiltonian in Eq.~\eqref{eq:drive_single} of the main text, we made the simplifying assumption that all drive amplitudes were real, which essentially enforces a phase-locking condition between the driving lasers.
+If the drive amplitudes are complex, $\Omega_m\to\Omega_me^{-\i\eta_m}$ (with real $\Omega_m,\eta_m$), then the single-spin drive Hamiltonian becomes
+\begin{multline}
+  \Delta \times H_{\t{3LD}}^{\t{single}}
+  = \p{\Omega_+^2 - \Omega_-^2} s_\z
+  - \sum_{\sigma\in\set{\pm1}} \Omega_0 \Omega_\sigma
+  \p{\sigma s_{\tilde\eta_\sigma,\x}
+    + s_\z s_{\tilde\eta_\sigma,\x}
+    + s_{\tilde\eta_\sigma,\x} s_\z} \\
+  - \Omega_0^2 s_\z^2 - \p{\Omega_+ + \Omega_-}^2 s_{\tilde\eta_0,\x}^2
+  - \p{\Omega_+ - \Omega_-}^2 s_{\tilde\eta_0,\y}^2,
+\end{multline}
+where $s_{\eta\alpha} \equiv e^{-\i\eta s_\z} s_\alpha e^{\i\eta s_\z}$ is a rotated spin-$\alpha$ operator (with e.g.~$s_{\pi/2,\x}=s_\y$), and
+\begin{align}
+  \tilde\eta_\pm \equiv \pm \p{\eta_\pm - \eta_0},
+  &&
+  \tilde\eta_0 \equiv \f{\eta_+ - \eta_-}{2}.
+\end{align}
 
@@
 \begin{align}
-  h_\crit^2 = \f{Q^2}{\eta^2}.
-\end{align}
+  h_\crit^2 \stackrel{?}{=} \f{Q^2}{\eta^2},
+\end{align}
+where we use the relation $\stackrel{?}{=}$ to indicate that this ``prediction'' of the Lax analysis is not necessarily valid for all $n$.
 For a permutationally symmetric state, up to vanishing $O(1/N)$ corrections we can expand
@@
 \begin{align}
-  h_\crit^2 = \f{n-1}{n} \times \f{12}{(n+1)n(n-1)}
+  h_\crit^2 \stackrel{?}{=} \f{n-1}{n} \times \f{12}{(n+1)n(n-1)}
   = \f{12}{n^2\p{n+1}}.

### 2021-04-21 sun_phases/spin_model_paper/main.tex
added conclusions, modified abstract and intro

@@
-\documentclass[nofootinbib,twocolumn]{revtex4-2}
+\documentclass[aps,pra,nofootinbib,twocolumn,superscriptaddress]{revtex4-2}
+\usepackage[T1]{fontenc}
 
@@
 
+% affiliations
+\newcommand{\JILA}{JILA, National Institute of Standards and Technology and University of Colorado, 440 UCB, Boulder, Colorado 80309, USA}
+\newcommand{\CTQM}{Center for Theory of Quantum Matter, University of Colorado, Boulder, CO, 80309, USA}
+\newcommand{\UO}{Homer L. Dodge Department of Physics and Astronomy, The University of Oklahoma, Norman, Oklahoma 73019, USA}
+\newcommand{\CQRT}{Center for Quantum Research and Technology, The University of Oklahoma, Norman, Oklahoma 73019, USA}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\title{Multilevel spin models in cold atomic systems}
+\title{Engineering an infinite-range SU($n$) spin model with ultracold fermions in an optical lattice}
 \author{Michael A. Perlin}
@@
 \author{Diego Barbarena}
+\affiliation{\JILA}
+\affiliation{\CTQM}
 \author{Robert J.~Lewis-Swan}
+\affiliation{\UO}
+\affiliation{\CQRT}
 % \orcid{https://orcid.org/0000-0002-0201-281X}
@@
 % \orcid{http://orcid.org/0000-0001-7176-9413}
-\affiliation{JILA, National Institute of Standards and Technology and University of Colorado, 440 UCB, Boulder, Colorado 80309, USA}
-\affiliation{Center for Theory of Quantum Matter, University of Colorado, Boulder, CO, 80309, USA}
+\affiliation{\JILA}
+\affiliation{\CTQM}
 
@@
 \begin{abstract}
-  Abstract.
+We study multilevel fermions in an optical lattice, with dynamics described by the Hubbard model with on site SU($n$)-symmetric interactions.
+We show that in an appropriate parameter regime, this system can be mapped onto a multilevel Heisenberg spin model with all-to-all SU($n$)-symmetric couplings.
+The use of Raman pulses to address internal spin states modifies the atomic dispersion relation and induces spin-orbit coupling, which can act as a synthetic inhomogeneous magnetic field that competes with the SU($n$) exchange interactions.
+We investigate the non-equilibrium mean-field dynamical phases of the resulting model as a function of $n$ and different initial configurations that are accessible with the Raman pulses.
+In agreement with previous studies with $n=2$, we find that for some initial states the spin model exhibits two distinct dynamical phases that obey simple scaling relations with $n$.
+When $n>2$, we find that dynamical behavior can be highly sensitive to initial intra-spin coherences.
+Our predictions are readily testable in current experiments with ultracold alkalike-earth(-like) atoms.
 \end{abstract}
@@
 
-The tremendous theoretical significance of SU($n$) symmetries makes it all the more exciting that they appear naturally in experimental atomic, molecular, and optical (AMO) platforms \cite{gorshkov2010twoorbital, beverland2016realizing, cazalilla2014ultracold, taie2012su, hofrichter2016direct, cappellini2014direct, scazza2014observation, zhang2014spectroscopic, goban2018emergence, perlin2019effective} with exquisite degrees of microscopic control.
-In the simplest example, this symmetry arises through the independence of atomic orbital and interaction parameters on the $n$ nuclear spin states of alkaline-earth(-like) atoms, with e.g.~$n=10$ in the case of ${}^{87}$Sr \cite{cazalilla2014ultracold}.
-As a result, experiments can directly probe the SU($n$) Hubbard model, leading to experimental observations of SU($n$) Hubbard phases and phase transitions \cite{taie2012su, hofrichter2016direct}, two-orbital SU($n$) magnetism \cite{cappellini2014direct, scazza2014observation, zhang2014spectroscopic}, and multi-body SU($n$)-symmetric interactions \cite{goban2018emergence, perlin2019effective}.
-In the spirit of quantum simulation, further investigations of this sort can play an important role in understanding SU($n$) symmetries' consequences for fundamental questions in physics, as well as their practical use in technological applications.
+The tremendous theoretical significance of SU($n$) symmetries makes it all the more exciting that they appear naturally in experimental atomic, molecular, and optical (AMO) platforms \cite{wu2003exact, cazalilla2009ultracold, gorshkov2010twoorbital, beverland2016realizing, cazalilla2014ultracold, taie2012su, hofrichter2016direct, cappellini2014direct, scazza2014observation, zhang2014spectroscopic, goban2018emergence, perlin2019effective} with exquisite degrees of microscopic control.
+In the simplest example, this symmetry arises through the independence of atomic orbital and interaction parameters on the $n$ nuclear spin states of alkaline-earth(-like) atoms, with e.g.~$n=10$ in the case of ${}^{87}$Sr \cite{wu2003exact, cazalilla2009ultracold, cazalilla2014ultracold}.
+As a result, experiments can directly probe the role SU($n$) interactions in controllable settings.
+Recent progress includes the experimental observations of bosonization and faster thermalization in multi-component Fermi gas with $SU(n)$ symmetry \cite{song2020evidence, sonderhouse2020thermodynamics}, SU($n$) Hubbard phases and phase transitions \cite{taie2012su, hofrichter2016direct, taie2020observation}, two-orbital SU($n$) magnetism \cite{cappellini2014direct, scazza2014observation, zhang2014spectroscopic}, and multi-body SU($n$)-symmetric interactions \cite{goban2018emergence, perlin2019effective}.
+In the spirit of quantum simulation, further investigations in controlled settings will play an important role in understanding SU($n$) symmetries' consequences for fundamental questions in physics, as well as their practical use in technological applications.
 For example, SU(2)-symmetric spin interactions can be harnessed to develop quantum sensors that surpass classical limits on measurement precision \cite{he2019engineering, perlin2020spin}.
-The exciting prospect of similarly exploiting larger SU($n$) symmetries to achieve a technological advantage is still an unexplored avenue of research with enormous potential.
-
-In this work, we ... \red{[finish intro, summarize paper]}
-% the (un)interested reader can ...
-% main results in/below Eqs. ...
+The prospect of similarly exploiting larger SU($n$) symmetries to achieve a technological advantage is still an unexplored avenue of research with untapped potential.
+
+In this work, we consider an experimentally relevant and theoretically tractable regime of the SU($n$) Hubbard model, highlighting differences and similarities with the more familiar case of SU(2).
+Working at ultracold temperatures and unit spatial filling (one atom per lattice site), we begin by mapping the SU($n$) Hubbard model onto a multilevel spin model with all-to-all SU($n$)-symmetric interactions in Section \ref{sec:spin_model}.
+We consider the effect of spin-orbit coupling (SOC) in Section \ref{sec:SOC}, finding in particular that the weak-SOC limit generally gives rise to a (synthetic) inhomogeneous magnetic field, extending previously known results to $n>2$ \cite{wall2016synthetic, livi2016synthetic, kolkowitz2016spinorbitcoupled, bromley2018dynamics, he2019engineering}.
+We take a look at control fields to address nuclear spins (and generate SOC) in Section \ref{sec:controls}, finding a simple three-laser driving scheme that allows for the preparation of interesting states with nontrivial intra-spin correlations when $n>2$.
+Putting all of these ingredients together, we examine mean-field dynamical behaviors of the SU($n$) spin model in Section \ref{sec:mean_field}, finding that
+\begin{enumerate*}
+\item long-time-averaged observables obey simple scaling relations with the spin dimension $n$, exhibiting (for spin-polarized initial states) a magnetized and a de-magnetized dynamical phase, as previously seen for the case of $n=2$ \cite{smale2019observation, lewis-swan2021cavityqed}, and
+\item when $n>2$ long-time dynamics can be highly sensitive to the intra-spin coherences of the initial state.
+\end{enumerate*}
+We conclude and discuss future outlooks in Section \ref{sec:conclusions}.
 
@@
 
-Before discussing SOC for $n$-level fermions, we first consider the well-studied case of two-level SOC on a one-dimensional lattice \cite{wall2016synthetic, bromley2018dynamics}.
+Before discussing SOC for $n$-level fermions, we first consider the well-studied case of two-level SOC on a one-dimensional lattice \cite{wall2016synthetic, kolkowitz2016spinorbitcoupled, bromley2018dynamics, he2019engineering}.
 In this case, SOC is induced by an external driving field that imprints a phase $e^{-\i\phi j}$ on lattice site $j$, or equivalently imparts a momentum kick $q\to q+\phi$, upon the absorption of a photon:
@@
 \begin{align}
-  \frac{\ket{s} + \ket{-s}}{\sqrt{2}}
-  \simeq e^{-\i\frac{\pi}{2}\p{s_\y + s_\y^2}} \ket{s},
-\end{align}
-where $\simeq$ denotes equality up to a global phase.
+  e^{-\i\frac{\pi}{2}\p{s_\y + s_\y^2}} \ket{s}
+  \propto \ket{s} + \ket{-s}.
+\end{align}
 This state has a vanishing mean spin vector, $\bk{s_\x}=\bk{s_\y}=\bk{s_\z}=0$, but variances $\bk{s_\x^2}=\bk{s_\y^2}=s/2$ and $\bk{s_\z^2}=s^2$.
@@
 \end{align}
-so
+where $\simeq$ denotes equality up to a global phase.
+The permutation $\Pi_\z$ commutes with $S_\y$, while the rotation $R_\x^\pi$ takes $S_\y\to-S_\y$, so
 \begin{align}
@@
 at all times.
-Conservation of $\bk{S_\y}=\bk{S_\z}=0$ and permutational symmetry together imply that the magnetization $\sigma_\MF$ is determined by the time-average of $\sigma_\x\equiv s_\x/s$ for a single spin:
+In turn, conservation of $\bk{S_\y}=\bk{S_\z}=0$ and permutational symmetry together imply that the magnetization $\sigma_\MF$ is determined by the time-average of $\sigma_\x\equiv s_\x/s$ for a single spin:
 \begin{align}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions}
-
-\red{[to be completed]}
+\section{Conclusions and outlooks}
+\label{sec:conclusions}
+
+Starting with an SU($n$) Hubbard model describing ultracold fermionic alkaline-earth(-like) atoms on an optical lattice, we derived a momentum-space multilevel spin model with all-to-all SU($n$)-symmetric interactions.
+We then considered the effect of spin-orbit coupling (SOC) induced by external driving lasers, finding that maintaining integrity of the spin model seems to require weak SOC, which in turn gives rise to a (synthetic) inhomogeneous magnetic field.
+Taking a closer look at the external controls that induce SOC, we found a simple three-laser drive that enables homogeneously addressing nuclear spins with a variety of spin Hamiltonians, enabling for example the preparation of ``multi-cat'' states in which each spin is in a coherent superposition of states spin-polarized in opposite directions.
+Finally, we examined dynamical behavior of the SU($n$) spin model at the mean-field level, finding that long-time observables obey simple scaling relations with $n$, and that when $n>2$ dynamical behavior can be highly sensitive to intra-spin coherences.
+
+Our work makes important progress in understanding the SU($n$) Fermi-Hubbard model in experimentally relevant parameter regimes, and we expect our findings to be readily testable in experiments with ultracold atoms.
+We hope this work provides a starting point for simulating SYK-like models in cold atomic platforms.
+In follow-up work, it would be interesting to study the the relationship between initial states and dynamical phases of our SU($n$) spin model more systematically, or to consider the effect of quantum corrections to mean-field behavior.
+There is also room to improve on the three-laser drive we described in this work, where it is natural to ask what ingredients and techniques are necessary to implement universal control of individual nuclear spins.
+Universal control would allow for an experimental study of $n$-dependence (including even/odd-$n$ parity effects) in a single experimental platform, simply by controlling the occupation and coherence of internal spin states.
+Finally, one can also study the SU($n$) Hubbard model in the super-exchange regime that gives rise to a real-space (as opposed to momentum-space) spin model, where SOC gives rise to chiral multilevel spin interactions.
+Unlike our present work, the super-exchange regime does not require weak SOC, and therefore has a larger parameter space in which to explore dynamical behaviors.
 
@@
 \end{align}
-or more generally an $M$-body operator\footnote{At face value, an $M$-body operator with $M>2$ does not typically appear in nature.
+or more generally an $M$-body operator\footnote{At face value, an $M$-body operator with $M>2$ does not typically appear in experiments.
 Nonetheless, considering $M>2$ illuminates the structure of eigenstates (and eigenvalues) of $H_0$, and allows us to go to high orders in perturbation theory with single- and two-body perturbations.}
@@
 \end{align}
-and if furthermore all $g_{ij}=-U/N$, as for $H_{\t{int}}$ in Eq.~\eqref{eq:H_int}, then the only relevant excitation energy is $\Delta=U$, and
+and if furthermore all $g_{ij}=-U/N$, as for $H_{\t{int}}$ in Eq.~\eqref{eq:H_int}, then the only relevant excitation energy is $\Delta=U$ (see Section \ref{sec:spin_wave}), and
 \begin{align}
@@
 \subsection{Recovering spin-wave theory}
+\label{sec:spin_wave}
 

### 2021-04-22 sun_phases/spin_model_paper/main.tex
updated paper in light of feedback from AMR and RLS

@@
 \renewcommand{\set}[1]{\left\{#1\right\}} % curly parenthesis
-\newcommand{\bk}{\Braket} % shorthand for braket notation
+\newcommand{\bk}{\braket} % shorthand for braket notation
+\newcommand{\Bk}{\Braket} % shorthand for braket notation
 \renewcommand{\v}{\bm} % bold vectors
-\renewcommand{\c}{\cdot} % inner product
 \renewcommand{\d}{\partial} % partial derivative
@@
 \newcommand{\1}{\mathds{1}}
+
+% symbols with hats
+\renewcommand{\b}{\hat b}
+\renewcommand{\c}{\hat c}
+\newcommand{\s}{\hat s}
+\renewcommand{\H}{\hat H}
+\renewcommand{\S}{\hat S}
+\newcommand{\R}{\hat R}
+\newcommand{\T}{\hat T}
+\renewcommand{\O}{\hat{\mathcal{O}}}
+\renewcommand{\P}{\hat{\mathcal{P}}}
 
@@
 \newcommand{\I}{\mathcal{I}}
-\renewcommand{\O}{\mathcal{O}}
-\renewcommand{\P}{\mathcal{P}}
 
@@
 \newcommand{\su}{\mathfrak{su}}
-\renewcommand{\ss}{\bar{\v s}\c\bar{\v s}}
+\renewcommand{\ss}{\bar{\v s}\cdot\bar{\v s}}
 \DeclareMathOperator{\diag}{diag}
@@
 \graphicspath{{./figures/}} % set path for all figures
-
-% for text markup
-\newcommand{\red}[1]{{\color{red} #1}}
 
@@
 We study multilevel fermions in an optical lattice, with dynamics described by the Hubbard model with on site SU($n$)-symmetric interactions.
-We show that in an appropriate parameter regime, this system can be mapped onto a multilevel Heisenberg spin model with all-to-all SU($n$)-symmetric couplings.
-The use of Raman pulses to address internal spin states modifies the atomic dispersion relation and induces spin-orbit coupling, which can act as a synthetic inhomogeneous magnetic field that competes with the SU($n$) exchange interactions.
-We investigate the non-equilibrium mean-field dynamical phases of the resulting model as a function of $n$ and different initial configurations that are accessible with the Raman pulses.
-In agreement with previous studies with $n=2$, we find that for some initial states the spin model exhibits two distinct dynamical phases that obey simple scaling relations with $n$.
-When $n>2$, we find that dynamical behavior can be highly sensitive to initial intra-spin coherences.
+We show that in an appropriate parameter regime this system can be mapped onto a multilevel spin model with all-to-all SU($n$)-symmetric couplings.
+Additionally, Raman pulses that address internal spin states modify the atomic dispersion relation and induce spin-orbit coupling, which can act as a synthetic inhomogeneous magnetic field that competes with the SU($n$) exchange interactions.
+We investigate the mean-field dynamical phase diagram of the resulting model as a function of $n$ and different initial configurations that are accessible with the Raman pulses.
+Consistent with previous studies for $n=2$, we find that for some initial states the spin model exhibits two distinct dynamical phases that obey simple scaling relations with $n$.
+Moreover, for $n>2$ we find that dynamical behavior can be highly sensitive to initial intra-spin coherences.
 Our predictions are readily testable in current experiments with ultracold alkalike-earth(-like) atoms.
@@
 Extensions of Yang-Mills and SU($n$) symmetry feature in the most well-studied examples of holographic duality \cite{maldacena1999largen} and the connection between entanglement and gravity \cite{ryu2006holographic} through the anti-de Sitter/conformal field theory (AdS/CFT) correspondence.
-In a condensed matter setting, SU(2) appears ubiquitously as a symmetry of the Hubbard model, with important consequences for the study of quantum magnetism and high temperature superconductivity \cite{lee2006doping}.
-The extension of SU(2) Hubbard and spin models to SU($n$) has led to predictions of exotic phases of matter such as valence bond solids \cite{read1989valencebond, rokhsar1990quadratic, kaul2012lattice, hermele2011topological} and chiral spin liquids \cite{hermele2009mott, hermele2011topological, chen2016syntheticgaugefield, nataf2016chiral}, as well as the potential to perform universal topological quantum computation \cite{freedman2004class, nayak2008nonabelian} and other phenomena \cite{nataf2014exact, nataf2016exact}.
-Furthermore, the consideration of disordered SU($n$) spin models has opened analytically tractable avenues for studying quantum chaos and information scrambling \cite{sachdev1993gapless, bentsen2019integrable}.
+In a condensed matter setting, SU($2$) appears ubiquitously as a symmetry of the Hubbard model, with important consequences for the study of quantum magnetism and high temperature superconductivity \cite{lee2006doping}.
+The extension of SU($2$) Hubbard and spin models to SU($n$) has led to predictions of exotic phases of matter such as valence bond solids \cite{read1989valencebond, rokhsar1990quadratic, kaul2012lattice, hermele2011topological} and chiral spin liquids \cite{hermele2009mott, hermele2011topological, chen2016syntheticgaugefield, nataf2016chiral}, as well as the potential to perform universal topological quantum computation \cite{freedman2004class, nayak2008nonabelian} and other phenomena \cite{nataf2014exact, nataf2016exact}.
+Furthermore, disordered SU($n$) spin models have opened analytically tractable avenues for studying quantum chaos and information scrambling \cite{sachdev1993gapless, bentsen2019integrable}.
 
 The tremendous theoretical significance of SU($n$) symmetries makes it all the more exciting that they appear naturally in experimental atomic, molecular, and optical (AMO) platforms \cite{wu2003exact, cazalilla2009ultracold, gorshkov2010twoorbital, beverland2016realizing, cazalilla2014ultracold, taie2012su, hofrichter2016direct, cappellini2014direct, scazza2014observation, zhang2014spectroscopic, goban2018emergence, perlin2019effective} with exquisite degrees of microscopic control.
-In the simplest example, this symmetry arises through the independence of atomic orbital and interaction parameters on the $n$ nuclear spin states of alkaline-earth(-like) atoms, with e.g.~$n=10$ in the case of ${}^{87}$Sr \cite{wu2003exact, cazalilla2009ultracold, cazalilla2014ultracold}.
-As a result, experiments can directly probe the role SU($n$) interactions in controllable settings.
+As a simple example, this symmetry arises through the independence of atomic orbital and interaction parameters on the $n$ nuclear spin states of alkaline-earth(-like) atoms, with e.g.~$n=10$ for ${}^{87}$Sr \cite{wu2003exact, cazalilla2009ultracold, cazalilla2014ultracold}.
+As a result, AMO experiments can directly probe the role of SU($n$) interactions in controllable settings.
 Recent progress includes the experimental observations of bosonization and faster thermalization in multi-component Fermi gas with $SU(n)$ symmetry \cite{song2020evidence, sonderhouse2020thermodynamics}, SU($n$) Hubbard phases and phase transitions \cite{taie2012su, hofrichter2016direct, taie2020observation}, two-orbital SU($n$) magnetism \cite{cappellini2014direct, scazza2014observation, zhang2014spectroscopic}, and multi-body SU($n$)-symmetric interactions \cite{goban2018emergence, perlin2019effective}.
-In the spirit of quantum simulation, further investigations in controlled settings will play an important role in understanding SU($n$) symmetries' consequences for fundamental questions in physics, as well as their practical use in technological applications.
-For example, SU(2)-symmetric spin interactions can be harnessed to develop quantum sensors that surpass classical limits on measurement precision \cite{he2019engineering, perlin2020spin}.
-The prospect of similarly exploiting larger SU($n$) symmetries to achieve a technological advantage is still an unexplored avenue of research with untapped potential.
-
-In this work, we consider an experimentally relevant and theoretically tractable regime of the SU($n$) Hubbard model, highlighting differences and similarities with the more familiar case of SU(2).
+In the spirit of quantum simulation, further investigations in controlled settings will play an important role in understanding the consequence of SU($n$) symmetries for fundamental questions in physics, as well as their practical use in technological applications.
+For example, SU($2$)-symmetric spin interactions can be harnessed to develop quantum sensors that surpass classical limits on measurement precision \cite{he2019engineering, perlin2020spin}.
+The prospect of similarly exploiting more general SU($n$) symmetries to achieve a technological advantage is still an unexplored avenue of research with untapped potential.
+
+In this work, we consider an experimentally relevant and theoretically tractable regime of the SU($n$) Hubbard model, highlighting differences and similarities with the more familiar case of SU($2$).
 Working at ultracold temperatures and unit spatial filling (one atom per lattice site), we begin by mapping the SU($n$) Hubbard model onto a multilevel spin model with all-to-all SU($n$)-symmetric interactions in Section \ref{sec:spin_model}.
 We consider the effect of spin-orbit coupling (SOC) in Section \ref{sec:SOC}, finding in particular that the weak-SOC limit generally gives rise to a (synthetic) inhomogeneous magnetic field, extending previously known results to $n>2$ \cite{wall2016synthetic, livi2016synthetic, kolkowitz2016spinorbitcoupled, bromley2018dynamics, he2019engineering}.
-We take a look at control fields to address nuclear spins (and generate SOC) in Section \ref{sec:controls}, finding a simple three-laser driving scheme that allows for the preparation of interesting states with nontrivial intra-spin correlations when $n>2$.
-Putting all of these ingredients together, we examine mean-field dynamical behaviors of the SU($n$) spin model in Section \ref{sec:mean_field}, finding that
+In Section \ref{sec:controls} we examine the use of control fields to address nuclear spins (and generate SOC), finding a simple three-laser driving scheme that allows for the preparation of interesting states with nontrivial intra-spin correlations when $n>2$.
+Finally, we combine these ingredients to examine mean-field dynamical behaviors of the SU($n$) spin model in Section \ref{sec:mean_field}, finding that:
 \begin{enumerate*}
-\item long-time-averaged observables obey simple scaling relations with the spin dimension $n$, exhibiting (for spin-polarized initial states) a magnetized and a de-magnetized dynamical phase, as previously seen for the case of $n=2$ \cite{smale2019observation, lewis-swan2021cavityqed}, and
-\item when $n>2$ long-time dynamics can be highly sensitive to the intra-spin coherences of the initial state.
+\item long-time-averaged observables obey simple scaling relations with the spin dimension $n$, exhibiting (for spin-polarized initial states) dynamical ferromagnetic and dynamical paramegnetic phases, as previously seen for the case of $n=2$ \cite{smale2019observation, lewis-swan2021cavityqed}, and
+\item for $n>2$ the long-time dynamics can be highly sensitive to the intra-spin coherences of the initial state.
 \end{enumerate*}
-We conclude and discuss future outlooks in Section \ref{sec:conclusions}.
+We conclude and discuss future directions in Section \ref{sec:conclusions}.
 
@@
 
-Here we review the realization of a collective SU($n$) spin model in a system of ultracold alkaline-earth(-like) atoms on an optical lattice.
-Without external driving fields, the evolution of such atoms in their electronic ground state is governed by single-body kinetic and two-body interaction Hamiltonians of the form
-\begin{align}
-  H_{\t{kin}}
-  &= -\f{J}{2} \sum_{\bk{j,j'},\mu} c_{j\mu}^\dag c_{j'\mu} + \t{h.c.}, \\
-  H_{\t{int}}
-  &= \f{U_0}{2} \sum_{j,\mu,\nu}
-  c_{j\mu}^\dag c_{j\mu} c_{j\nu}^\dag c_{j\nu},
-\end{align}
-where $\bk{j,j'}$ denotes neighboring lattice sites $j$ and $j'$; $\mu,\nu\in\set{s,s-1,\cdots,-s}$ index orthogonal spin states of a spin-$s$ nucleus, with $s=\frac{n-1}{2}$ (e.g.~$s=\frac{9}{2}$ in the case of ${}^{87}$Sr with $10$ nuclear spin states); $c_{j\mu}$ is a fermionic annihilation operator, $J$ is a tunneling amplitude; and $U_0$ is a two-body on-site interaction energy.
-In the present work, we neglect inter-site interactions and interaction-assisted hopping, which may become relevant in a sufficiently shallow lattice, namely when $J\gtrsim E_{\t{R}}$, where $E_{\t{R}}$ is the atomic lattice recoil energy.
-For simplicity, we now assume a periodic lattice of $L$ sites, and expand the on-site fermionic operators in terms of operators addressing (quasi-)momentum modes $q$, $c_{j\mu}=\frac1{\sqrt{L}}\sum_k e^{-\i q\c j} c_{q\mu}$, finding that
-\begin{align}
-  H_{\t{kin}}
-  &= -J\sum_{q,\mu} \cos\p{q} c_{q\mu}^\dag c_{q\mu}, \\
-  H_{\t{int}}
-  &= \f{U}{2N} \sum_{k,\ell,p,q,\mu,\nu}
-  c_{k\mu}^\dag c_{\ell\mu} c_{p\nu}^\dag c_{q\nu}
+Here we review the realization of a collective SU($n$) spin model in a system of ultracold alkaline-earth(-like) atoms trapped in an optical lattice.
+Without external driving fields, the evolution of such atoms in their electronic ground state is governed by the single-body kinetic and two-body interaction Hamiltonians
+\begin{align}
+  \H_{\t{kin}}
+  &= -J \sum_{\bk{j,j'},\mu} \c_{j\mu}^\dag \c_{j'\mu} + \t{h.c.}, \\
+  \H_{\t{int}}
+  &= \f{U}{2} \sum_{j,\mu,\nu}
+  \c_{j\mu}^\dag \c_{j\mu} \c_{j\nu}^\dag \c_{j\nu},
+\end{align}
+where $\bk{j,j'}$ denotes neighboring lattice sites $j$ and $j'$; $\mu,\nu\in\set{s,s-1,\cdots,-s}$ index orthogonal spin states of a spin-$s$ nucleus, with $s=\frac{n-1}{2}$ (e.g.~$s=\frac{9}{2}$ in the case of ${}^{87}$Sr with $10$ nuclear spin states); $\c_{j\mu}$ is a fermionic annihilation operator, $J$ is a tunneling amplitude (for simplicity assumed to be the same in all directions); and $U$ is a two-body on-site interaction energy.
+In the present work, we neglect inter-site interactions and interaction-assisted hopping, which is a good approximation for a sufficiently deep lattice, namely when $J\lesssim E_{\t{R}}$, where $E_{\t{R}}$ is the atom recoil energy.
+For simplicity, we now assume a one-dimensional periodic lattice of $L$ sites, and expand the on-site fermionic operators in terms of operators addressing (quasi-)momentum modes $q$ (in units with lattice spacing $a=1$), $\c_{j\mu} = \frac1{\sqrt{L}} \sum_k e^{-\i q\cdot j} \c_{q\mu}$, finding that
+\begin{align}
+  \H_{\t{kin}}
+  &= -2J\sum_{q,\mu} \cos\p{q} \c_{q\mu}^\dag \c_{q\mu}, \\
+  \H_{\t{int}}
+  &= \f{u}{2N} \sum_{k,\ell,p,q,\mu,\nu}
+  \c_{k\mu}^\dag \c_{\ell\mu} \c_{p\nu}^\dag \c_{q\nu}
   \times \delta_{k+p,\ell+q},
@@
 \end{align}
-where $N$ is the total number of atoms on the lattice, we define $U\equiv U_0\times N/L$ for convenience, $\delta_{k+p,\ell+q}=1$ if $k+p=\ell+q$ and zero otherwise (enforcing conservation of momentum), and we work in units with lattice spacing $a=1$.
-
-If the interaction energy $U_0$ is smaller than the single-particle bandwidth $\sim J$, then the mode-changing collisions in $H_{\t{int}}$ become off-resonant, motivating the frozen-mode approximation $\set{k,p}=\set{\ell,q}$ (which we assume for the remainder of this work)\footnote{Note that the frozen-mode approximation neglects correlated momentum-hopping terms of the form $c_{\pi-p,\mu}^\dag c_{\pi-q,\mu} c_{p\nu}^\dag c_{q\nu}$, which conserve both momentum and energy.
+where $N$ is the total number of atoms on the lattice, we define $u\equiv U\times N/L$ for convenience, $\delta_{k+p,\ell+q}=1$ if $k+p=\ell+q$ and zero otherwise (enforcing conservation of momentum).
+
+If the interaction energy $U$ is small compared to the single-particle bandwidth $4J$, then the mode-changing collisions in $\H_{\t{int}}$ become off-resonant, motivating the frozen-mode approximation $\set{k,p}=\set{\ell,q}$ (which we assume for the remainder of this work)\footnote{Note that the frozen-mode approximation neglects correlated momentum-hopping terms of the form $\c_{\pi-p,\mu}^\dag \c_{\pi-q,\mu} \c_{p\nu}^\dag \c_{q\nu}$, which conserve both momentum and energy.
 We defer a careful treatment of these terms to future work, noting only that they vanish on the manifold of permutationally symmetric spin states with one atom per lattice site.}.
-The operator content of terms with $k=\ell$ and $p=q$ takes the form $c_{\ell\mu}^\dag c_{\ell\mu} c_{q\nu}^\dag c_{q\nu}$, so these terms contribute an overall energy $\frac12NU$ that can be neglected in the absence of coherent atom number fluctuations.
-Defining the spin operators $s_{\mu\nu q}\equiv c_{q\mu}^\dag c_{q\nu}$, the remaining terms of the kinetic and interaction Hamiltonians become
-\begin{align}
-  H_{\t{kin}} &= -J\sum_{q,\mu} \cos\p{q} s_{\mu\mu q},
+The terms with $k=\ell$ and $p=q$ are $\frac{u}{2N} \sum \c_{\ell\mu}^\dag \c_{\ell\mu} \c_{q\nu}^\dag \c_{q\nu} = \frac12 N u$, which is an overall energy shift that we can freely neglect.
+Defining the spin operators $\s_{\mu\nu q}\equiv \c_{q\mu}^\dag \c_{q\nu}$, the remaining terms of the kinetic and interaction Hamiltonians are
+\begin{align}
+  \H_{\t{kin}}
+  &= -2J\sum_{q,\mu} \cos\p{q} \s_{\mu\mu q},
   \label{H_kin_start} \\
-  H_{\t{int}} &= -\f{U}{2N} \sum_{p,q,\mu,\nu} s_{\mu\nu p} s_{\nu\mu q}.
+  \H_{\t{int}}
+  &= -\f{u}{2N} \sum_{p,q,\mu,\nu} \s_{\mu\nu p} \s_{\nu\mu q}.
   \label{eq:H_int_start}
@@
 Throughout this work, we will assume that atomic modes are singly-occupied, e.g.~due to the initialization of a spin-polarized state with one atom per lattice site, in which multiple occupation of an atomic mode is forbidden by fermionic statistics (Pauli exclusion).
-In this case, we can neglect the fermionic origins of $H_{\t{int}}$ and instead treat our system as a bona fide spin model.
-Note that the ``kinetic'' terms of this spin model ($H_{\t{kin}}$) are proportional to the identity operator, and can be neglected for the time being.
-The fermionic origin of these terms will return to play an important role when we consider spin-orbit coupling in Section \ref{sec:SOC}.
-
-To further simplify the interaction Hamiltonian $H_{\t{int}}$ and write it in a form suggestively similar to SU(2) spin alignment, we now construct the operator-valued spin matrix
-\begin{align}
-  \v s_q \equiv \sum_{\mu,\nu} s_{\mu\nu q} \op{\mu}{\nu},
-\end{align}
-and for any pair of such operator-valued matrices $\v A,\v B$, we define the inner product
-\begin{align}
-  \v A \c \v B \equiv \Tr_{\t{aux}}\p{\v A^\dag\v B}
-  = \sum_{\mu,\nu} A_{\mu\nu}^\dag B_{\mu\nu},
+In this case we can simply treat our system as $N$ distinguishable $n$-level quantum spins at ``lattice sites'' $p,q$.
+Note that the ``kinetic'' terms of this spin model ($\H_{\t{kin}}$) are proportional to the identity operator, contributing an overall shift in energy that we can neglect and this point.
+Nevertheless, these kinetic terms will become important in the presence of an external drive, which we discuss in Section \ref{sec:SOC}.
+
+To further simplify the interaction Hamiltonian $\H_{\t{int}}$ and write it in a form reminiscent of more familiar SU($2$) spin models, we now construct the operator-valued spin matrix
+\begin{align}
+  \v\s_q \equiv \sum_{\mu,\nu} \s_{\mu\nu q} \op{\mu}{\nu},
+\end{align}
+and for any pair of such operator-valued matrices $\v{\hat{A}},\v{\hat{B}}$, we define the inner product
+\begin{align}
+  \v{\hat{A}} \cdot \v{\hat{B}}
+  \equiv \sum_{\mu,\nu} A_{\mu\nu}^\dag B_{\mu\nu}.
   \label{eq:dot_product}
 \end{align}
-where the trace is performed over the auxiliary matrix degrees of freedom.
 These definitions allow us to write the spin Hamiltonian in Eq.~\eqref{eq:H_int_start} as
 \begin{align}
-  H_{\t{int}} = -\f{U}{2N} \sum_{p,q} \v s_p\c\v s_q
-  = -\f{U}{2N}\v S\c\v S,
-  &&
-  \v S \equiv \sum_q \v s_q.
+  \H_{\t{int}} = -\f{u}{2N} \sum_{p,q} \v\s_p\cdot\v\s_q
+  = -\f{u}{2N}\v\S\cdot\v\S,
+  &&
+  \v\S \equiv \sum_q \v\s_q.
   \label{eq:H_int}
 \end{align}
-Here $\v S$ is a collective spin matrix, analogous to the collective spin vector $\vec S=\p{S_\x,S_\y,S_\z}$ in the case of SU(2) \cite{he2019engineering}, with $\v S\c\v S \simeq 2 \vec S\c\vec S = 2 \sum_{\alpha\in\set{\x,\y,\z}}S_\alpha^2$ when $n=2$ (here $\simeq$ denotes equality after modding out identity terms).
+Here $\v\S$ is a collective spin matrix, analogous to the collective spin vector $\vec S=(\S_\x,\S_\y,\S_\z)$ in the case of SU($2$) \cite{he2019engineering}, with $\frac12\v\S\cdot\v\S \simeq \vec S\cdot\vec S = \S_\x^2 + \S_\y^2 + \S_\z^2$ when $n=2$ (here $\simeq$ denotes equality after modding out identity terms).
 
@@
 \caption{
-(a) Ultracold atoms on a lattice of $L$ sites can tunnel between neighboring lattice sites at a rate $J$, and locally repel each other with interaction energy $U_0$.
-(b,c) When the interaction energy $U_0$ is smaller than the single-particle bandwidth $\sim J$, the frozen-mode approximation enables writing the interaction Hamiltonian as a spin model consisting of exchange terms $\v s_p\c\v s_q$, which swap the states two spins pinned to momentum modes $p,q$.
-(d) Interactions thereby open an energy gap $U=U_0\times N/L$ between the manifold of permutationally symmetric states of $N$ spins, and the orthogonal complement of states that break spin-permutation symmetry.
+(a) Ultracold atoms on a lattice of $L$ sites tunnel between neighboring lattice sites at a rate $J$, and locally repel each other with interaction energy $U$.
+(b,c) When the interaction energy $U$ is small compared to the single-particle bandwidth $4J$, the frozen-mode approximation enables the interaction Hamiltonian to be written as a spin model consisting of exchange terms $\v\s_p\cdot\v\s_q$, which swap the states two spins pinned to momentum modes $p,q$.
+(d) Interactions open an energy gap $u=U\times N/L$ between the manifold of permutationally symmetric states of $N$ spins, and the orthogonal complement of states that break spin-permutation symmetry.
 }
@@
 
-We now discuss the spin Hamiltonian $H_{\t{int}}$ in Eq.~\eqref{eq:H_int}.
-The operator $\v s_p\c\v s_q$ simply swaps the nuclear spin states of two atoms pinned to modes $p,q$.
-The term $-\v s_p\c\v s_q$ thereby assigns a definite energy of $-1$ ($+1$) to a pair of spins that are symmetric (anti-symmetric) under exchange.
-In this sense, $\v s_p\c\v s_q$ is analogous to the enforcement of SU(2) spin alignment by ferromagnetic interactions, which similarly assigns (distinct) definite energies to anti-symmetric spin-0 singlet $\ket{\up\dn}-\ket{\dn\up}$ and the symmetric spin-1 triplets $\set{\ket{\up\up},\ket{\dn\dn},\ket{\up\dn}+\ket{\dn\up}}$.
-By summing over all pair-wise exchange terms $\v s_p\c\v s_q$, the interaction Hamiltonian $H_{\t{int}}$ energetically enforces a permutational symmetry among all spins, opening an energy gap $U$ between the manifold of all permutationally symmetric (PS) states and the orthogonal complement of excited (e.g.~spin-wave) states that break permutational symmetry.
+We now discuss the spin Hamiltonian $\H_{\t{int}}$ in Eq.~\eqref{eq:H_int}.
+The operator $\v\s_p\cdot\v\s_q$ simply swaps the nuclear spin states of two atoms pinned to modes $p,q$.
+The term $-\v\s_p\cdot\v\s_q$ thereby assigns a definite energy of $-1$ ($+1$) to a pair of spins that are symmetric (anti-symmetric) under exchange.
+In this sense, $\v\s_p\cdot\v\s_q$ is analogous to the enforcement of SU($2$) spin alignment by ferromagnetic interactions, which similarly assigns different energies to the anti-symmetric spin-0 singlet $\ket{\up\dn}-\ket{\dn\up}$ and the symmetric spin-1 triplets $\set{\ket{\up\up},\ket{\dn\dn},\ket{\up\dn}+\ket{\dn\up}}$.
+By summing over all pair-wise exchange terms $\v\s_p\cdot\v\s_q$, the interaction Hamiltonian $\H_{\t{int}}$ energetically enforces a permutational symmetry among all spins, opening an energy gap $u$ between the manifold of all permutationally symmetric (PS) states and the orthogonal complement of excited (e.g.~spin-wave) states that break permutational symmetry.
 See Figure \ref{fig:spin_model} for a summary of this section thus far.
 
-In the case of SU(2), the PS manifold is precisely the Dicke manifold of collective states $\ket{m_\z}$ with total spin $S=\frac{N}{2}$ and definite spin projection $m_\z\in\set{S,S-1,\cdots,-S}$ onto a fixed quantization axis.
+In the case of SU($2$), the PS manifold is precisely the Dicke manifold of collective states $\ket{m_\z}$ with total spin $S=\frac{N}{2}$ and definite spin projection $m_\z\in\set{S,S-1,\cdots,-S}$ onto a fixed quantization axis.
 Equivalently, Dicke states $\ket{m_\z}=\ket{m_\up,m_\dn}$ can be labeled by a definite number of spins $m_\up=S+m_\z$ ($m_\dn=S-m_\z$) pointing up (down) along the spin quantization axis, with $m_\up+m_\dn=N$.
@@
 External fields or additional interactions that respect permutational symmetry can induce nontrivial dynamics within the PS manifold.
-Even if additional Hamiltonian terms break permutational symmetry, as long as the coupling to states outside of the PS manifold is small compared to the interaction energy $U$, these terms can be restricted to the PS manifold perturbatively (see Appendix \ref{sec:pert_theory}) \cite{bravyi2011schrieffer}.
+Moreover, additional terms that explicitly break permutational symmetry can nevertheless lead to interesting dynamics that can be captured within the PS manifold perturbatively, as long as the coupling to non-PS states is weak compared $u$ (see Appendix \ref{sec:pert_theory}) \cite{bravyi2011schrieffer}.
+This perturbative regime is thereby efficiently simulable, as the PS manifold has dimension $\sim N^{n-1}$ (as compared to $n^N$ for the entire spin Hilbert space).
 Simulating dynamics within the PS manifold requires calculating matrix elements $\bk{\ell|\O|m}$ of spin operators $\O$ with respect to PS states $\ket\ell,\ket{m}$; we discuss this calculation in Appendix \ref{sec:PS_ops}.
@@
 Finally, we take a moment to discuss individual $n$-level spins.
-Whenever possible, it is desirable to find geometric representations of abstract mathematical objects, such as the state of a quantum system.
-Geometric representations facilitate the development of understanding and intuitions that can be difficult to extract from algebraic expressions alone.
-For this reason, the state of a two-level spin, or a qubit, is commonly represented by a point on (or within) the Bloch sphere.
-More generally, the state $\ket\psi$ of a $n$-level spin can be represented by a probability distribution\footnote{Strictly speaking, in order to make $Q_\psi$ (and $Q_\rho$) a probability distribution, one has to divide it by a normalization factor $\frac{4\pi}{d}$.
+The state of a two-level spin, or a qubit, is commonly represented by a point on (or within) the Bloch sphere.
+More generally, the state $\ket\psi$ of a $n$-level spin can be represented by a probability distribution\footnote{Strictly speaking, in order to make $Q_\psi$ (and $Q_{\hat\rho}$) a probability distribution, one has to divide it by a normalization factor $\frac{4\pi}{d}$.
 We omit this normalization factor for convenience.} $Q_\psi$ on the sphere $\SS^2$.
 The value $Q_\psi\p{\v v}$ at a point $\v v\in\SS^2$ equal to the overlap of $\ket\psi$ with a pure state $\ket{\v v}$ maximally polarized along $\v v$: $Q_\psi\equiv\abs{\bk{\v v|\psi}}^2$ (see Figure \ref{fig:spin_dist}).
-In the case of a mixed state $\rho$, this distribution is defined by $Q_\rho\equiv\bk{\v v|\rho|\v v}$.
-The dimension $n$ is reflected in $Q_\rho$ by the fact that it can be decomposed into spherical harmonics $Y_{\ell m}$ with degree $\ell<n$ (that is, $Q_\rho$ is ``low resolution'' or ``band-limited'' at degree $n$) \cite{perlin2020qudit}, as well as in the normalization $\int_{\v v\in\SS^2}\dd^2\v v\,Q_\rho\p{\v v}=\frac{4\pi}{d}$.
+In the case of a mixed state $\hat\rho$, this distribution is defined by $Q_{\hat\rho}\equiv\bk{\v v|\hat\rho|\v v}$\footnote{The dimension $n$ is reflected in $Q_{\hat\rho}$ by the fact that it can be decomposed into spherical harmonics $Y_{\ell m}$ with degree $\ell<n$ (that is, $Q_{\hat\rho}$ is ``low resolution'' or ``band-limited'' at degree $n$) \cite{perlin2020qudit}, as well as in the normalization $\int_{\v v\in\SS^2}\dd^2\v v\,Q_{\hat\rho}\p{\v v}=\frac{4\pi}{d}$.}.
 Closely related spherical representations of multilevel spin states (and operators) are discussed in Ref.~\cite{li2013weylwignermoyal}.
@@
 
-In fact, $Q_\rho$ is essentially the Husimi-$Q$ function commonly used (e.g.~in the spin squeezing community \cite{ma2011quantum}) to represent collective states of spin-$\frac{1}{2}$ particles in the Dicke manifold.
+In fact, $Q_{\hat\rho}$ is essentially the Husimi-$Q$ function commonly used (e.g.~in the spin squeezing community \cite{ma2011quantum}) to represent collective states of spin-$\frac{1}{2}$ particles in the Dicke manifold.
 This correspondence is made precise by identifying the Hilbert space of a single $n$-level spin with the Dicke manifold of $n-1$ spin-$\frac{1}{2}$ particles.
-The Hilbert space of a 10-level (spin-$\frac92$) nucleus of a ${}^{87}$Sr atom, for example, can be thought of as the Dicke manifold of 9 ``excess'' proton and neutron spins.
+The Hilbert space of a 10-level (spin-$\frac92$) nucleus of a ${}^{87}$Sr atom, for example, can be thought of as the Dicke manifold of 9 spin-$\frac{1}{2}$ particles (``excess'' protons and neutrons).
+We will freely use this correspondence to import standard results and constructions for the Dicke manifold (e.g.~the construction of spin operators $\s_\x,\s_\y,\s_\z$) into the context of a multilevel spin.
 
@@
 
-We wish to study the dynamical behaviors of experimentally realizable SU($n$) spin models.
-In order to ensure validity of the $\v S\c\v S$ spin Hamiltonian in Eq.~\eqref{eq:H_int}, we consider initial preparation of a spin-polarized state\footnote{We remind the reader that validity of the spin model requires that each momentum mode is at most singly-occupied, which is guaranteed for a spin-polarized state due to Pauli blocking.}, which has the added benefit of being simple to prepare experimentally.
-However, a spin polarized state is an eigenstate of the $\v S\c\v S$ spin Hamiltonian, so we need additional ingredients for nontrivial behavior.
+We wish to study the dynamical behavior of experimentally realizable SU($n$) spin models.
+In order to ensure validity of the $\v\S\cdot\v\S$ spin Hamiltonian in Eq.~\eqref{eq:H_int}, we consider initial preparation of a spin-polarized state\footnote{We remind the reader that validity of the spin model requires that each momentum mode is at most singly-occupied, which is guaranteed for a spin-polarized state due to Pauli blocking.}, which has the added benefit of being simple to prepare experimentally.
+However, a spin polarized state is an eigenstate of the $\v\S\cdot\v\S$ spin Hamiltonian, so we need additional ingredients for nontrivial behavior.
 Here, we consider the addition of spin-orbit coupling (SOC) induced by external driving fields.
 
-Before discussing SOC for $n$-level fermions, we first consider the well-studied case of two-level SOC on a one-dimensional lattice \cite{wall2016synthetic, kolkowitz2016spinorbitcoupled, bromley2018dynamics, he2019engineering}.
-In this case, SOC is induced by an external driving field that imprints a phase $e^{-\i\phi j}$ on lattice site $j$, or equivalently imparts a momentum kick $q\to q+\phi$, upon the absorption of a photon:
-\begin{align}
-  H_{\t{drive}}^{(\phi)}
-  = \f{\Omega}{2} \sum_q c_{q+\phi,\up}^\dag c_{q,\dn} + \t{h.c.}.
+Before discussing SOC for $n$-level fermions, we first consider the well-studied case of two-level SOC with a one-dimensional lattice \cite{wall2016synthetic, kolkowitz2016spinorbitcoupled, bromley2018dynamics, he2019engineering}.
+In this case, SOC is induced by an external driving field that imprints a phase $e^{-\i\phi j}$ on lattice site $j$, or equivalently imparts a momentum kick $q\to q+\phi$, upon the absorption of a photon\footnote{In order for the drive Hamiltonian $\H_{\t{drive}}^{(\phi)}$ to be well-defined, $\phi$ should be commensurate with the lattice, with e.g.~$\phi\in2\pi/L\times\ZZ$ on a one-dimensional lattice of $L$ sites.}:
+\begin{align}
+  \H_{\t{drive}}^{(\phi)}
+  = \f{\Omega}{2} \sum_q \c_{q+\phi,\up}^\dag \c_{q,\dn} + \t{h.c.}.
   \label{eq:drive_2}
 \end{align}
-Identifying a numerical spin index $\mu=+\frac12$ ($-\frac12$) with the state $\up$ ($\dn$), this drive Hamiltonian can be diagonalized in its momentum index $q$ by the gauge transformation $c_{q\mu}^\dag\to c_{q-\mu\phi,\mu}^\dag$ (equivalently $c_{j\mu}^\dag\to e^{\i\phi\mu j} c_{j\mu}^\dag$), which takes
-\begin{align}
-  H_{\t{drive}}^{(\phi)} \to H_{\t{drive}} \equiv \Omega S_\x,
-  &&
-  S_\x \equiv \sum_q s_{\x,q},
+Identifying a numerical spin index $\mu=+\frac12$ ($-\frac12$) with the state $\up$ ($\dn$), this drive Hamiltonian can be diagonalized in its momentum index $q$ by the gauge transformation $\c_{q\mu}^\dag\to \c_{q-\mu\phi,\mu}^\dag$ (equivalently $\c_{j\mu}^\dag\to e^{\i\phi\mu j} \c_{j\mu}^\dag$), which takes
+\begin{align}
+  \H_{\t{drive}}^{(\phi)} \to \H_{\t{drive}} \equiv \Omega \S_\x,
+  &&
+  \S_\x \equiv \sum_q \s_{\x,q},
   \label{eq:drive_trans}
 \end{align}
-where $S_\x$ is a collective spin-$x$ operator, and $s_{\x,q}=\frac12 c_{q,\up}^\dag c_{q,\dn} + \t{h.c.}$ for two-level spins.
-
-The two-level drive in Eq.~\eqref{eq:drive_2} is typically implemented by coupling pseudo-spin electronic states of an atom with an external laser \cite{wall2016synthetic, bromley2018dynamics}.
-Though the coupling between laser fields and nuclear spins is generally too weak to drive nuclear spin transitions directly, it will nonetheless be useful to consider the $n$-level generalization of the two-level drive.
-This generalization will capture the important features of SOC for $n$-level nuclei, while avoiding complications that we defer to Section \ref{sec:controls}.
+where $\S_\x$ is a collective spin-x operator, and $\s_{\x,q}=\frac12 \c_{q,\up}^\dag \c_{q,\dn} + \t{h.c.}$ for two-level spins.
+
+The two-level SOC drive in Eq.~\eqref{eq:drive_2} is typically implemented by coupling two electronic states of an atom with an external laser \cite{wall2016synthetic, livi2016synthetic, kolkowitz2016spinorbitcoupled, bromley2018dynamics, he2019engineering}.
+Though the coupling between optical laser fields and nuclear spins is generally too weak to drive nuclear spin transitions directly, it will nonetheless be useful to consider the $n$-level generalization of the two-level drive.
+This generalization will capture the important features of SOC for $n$-level spins, while avoiding complications that we defer to Section \ref{sec:controls}.
 
@@
 \begin{align}
-  H_{\t{drive}}^{(\phi)} = \Omega_0 \sum_{q,\mu} \bk{s,\mu;1,1|s,\mu+1}
-  c_{q+\phi,\mu+1}^\dag c_{q\mu} + \t{h.c.},
+  \H_{\t{drive}}^{(\phi)} = \Omega_0 \sum_{q,\mu} \bk{s,\mu;1,1|s,\mu+1}
+  \c_{q+\phi,\mu+1}^\dag \c_{q\mu} + \t{h.c.},
 \end{align}
 where $\mu\in\set{s,s-1,\cdots,-s}$, and $\bk{s,\mu;1,1|s,\mu+1}$ is a Clebsch-Gordan coefficient.
-The gauge gauge transformation $c_{q\mu}^\dag\to c_{q-\mu\phi,\mu}^\dag$ again takes $H_{\t{drive}}^{(\phi)}\to \Omega S_\x$ as in Eq.~\eqref{eq:drive_trans}, where now
+The gauge gauge transformation $\c_{q\mu}^\dag\to \c_{q-\mu\phi,\mu}^\dag$ again takes $\H_{\t{drive}}^{(\phi)}\to \Omega \S_\x$ as in Eq.~\eqref{eq:drive_trans}, where now
 \begin{align}
@@
 \begin{align}
-  s_{\x,q} \equiv \xi \sum_\mu \bk{s,\mu;1,1|s,\mu+1} c_{q,\mu+1}^\dag c_{q\mu} + \t{h.c.}
+  \s_{\x,q} \equiv \xi \sum_\mu \bk{s,\mu;1,1|s,\mu+1} \c_{q,\mu+1}^\dag \c_{q\mu} + \t{h.c.}
   \label{eq:spin_x_op}
 \end{align}
-is an $n$-level spin-$x$ operator for spin $q$, with which $e^{-\i\theta s_{\x,q}}$ rotates the state of spin $q$ about the $x$ axis by an angle $\theta$ (literally rotating the sphere in Figure \ref{fig:spin_dist}).
-Despite complicating details, the change from two-level to $n$-level spins thus results in a drive Hamiltonian with an identical physical interpretation: after moving into an appropriate ``gauged frame'', this drive simply rotates all spins about the $x$ axis at a rate $\Omega$.
+is an $n$-level spin-x operator for spin $q$, with which $e^{-\i\theta \s_{\x,q}}$ rotates the state of spin $q$ about the x axis by an angle $\theta$ (literally rotating the sphere in Figure \ref{fig:spin_dist}).
+Despite complicating details, the change from two-level to $n$-level spins thus results in a drive Hamiltonian with an identical physical interpretation: after moving into an appropriate ``gauged frame'', this drive simply rotates all spins about the x axis at a rate $\Omega$.
 
@@
 \begin{align}
-  H_{\t{kin}} \to H_{\t{kin}}^{(\phi)}
-  \equiv -J \sum_q \cos\p{q+\mu\phi} s_{\mu\mu q},
+  \H_{\t{kin}} \to \H_{\t{kin}}^{(\phi)}
+  \equiv -2J \sum_q \cos\p{q+\mu\phi} \s_{\mu\mu q},
 \end{align}
@@
 \begin{align}
-  H_{\t{kin}}^{(\phi)}
-  = -J \sum_q
-  \sp{\cos\p{q} w_{+,q}^{(\phi)} - \sin\p{q} w_{+,q}^{(\phi)}},
+  \H_{\t{kin}}^{(\phi)}
+  = -2J \sum_q
+  \sp{\cos\p{q} \hat w_{+,q}^{(\phi)} - \sin\p{q} \hat w_{+,q}^{(\phi)}},
 \end{align}
@@
 \begin{align}
-  w_{+,q}^{(\phi)} &\equiv \sum_\mu \cos\p{\mu\phi} s_{\mu\mu q}, \\
-  w_{-,q}^{(\phi)} &\equiv \sum_\mu \sin\p{\mu\phi} s_{\mu\mu q}.
-\end{align}
-For two-level spins, $w_{+,q}^{(\phi)}$ is proportional to the identity operator, and $w_{+,q}^{(\phi)}=2\sin\p{\phi/2} s_{\z,q}$, where $s_{\z,q}=\sum_\mu \mu\, s_{\mu\mu q}$ is a spin-$z$ operator for spin $q$, so the kinetic Hamiltonian in the gauged frame describes a (synthetic) inhomogeneous magnetic field:
-\begin{align}
-  \left. H_{\t{kin}}^{(\phi)} \right|_{n=2}
-  = 2\sin\p{\phi/2} J \sum_q \sin\p{q} s_{\z,q}.
+  \hat w_{+,q}^{(\phi)} &\equiv \sum_\mu \cos\p{\mu\phi} \s_{\mu\mu q}, \\
+  \hat w_{-,q}^{(\phi)} &\equiv \sum_\mu \sin\p{\mu\phi} \s_{\mu\mu q}.
+\end{align}
+For two-level spins with $\mu=\pm\frac12$, $\hat w_{+,q}^{(\phi)}$ is proportional to the identity operator, and $\hat w_{+,q}^{(\phi)}=2\sin\p{\phi/2} \s_{\z,q}$, where $\s_{\z,q}=\sum_\mu \mu\, \s_{\mu\mu q}$ is a spin-z operator for spin $q$, so the kinetic Hamiltonian in the gauged frame describes a (synthetic) inhomogeneous magnetic field:
+\begin{align}
+  \left. \H_{\t{kin}}^{(\phi)} \right|_{n=2}
+  = 4\sin\p{\phi/2} J \sum_q \sin\p{q} \s_{\z,q}.
 \end{align}
@@
 \begin{align}
-  \left. H_{\t{kin}}^{(\phi)} \right|_{s\phi\ll1}
-  = J\phi \sum_q \sin\p{q} s_{\z,q} + O\p{(s\phi)^2}.
-\end{align}
-As the strength of SOC is increased, this Hamiltonian acquires terms with higher powers of $s_{\z,q}$, up to $s_{\z,q}^{n-1}$.
-
-In principle, making the gauge transformation $c_{q\mu}^\dag\to c_{q-\mu\phi,\mu}^\dag$ also transforms the interaction Hamiltonian as $H_{\t{int}} \to H_{\t{int}}^{(\phi)}$, where
-\begin{align}
-  H_{\t{int}}^{(\phi)}
-  \equiv -\f{U}{2N} \sum_{p,q,\mu,\nu}
-  c_{p-\mu\phi,\mu}^\dag c_{p-\nu\phi,\nu}
-  c_{q-\nu\phi,\nu}^\dag c_{q-\mu\phi,\mu}.
-\end{align}
-However, $H_{\t{int}}^{(\phi)}$ cannot be exactly reduced to a spin Hamiltonian that addresses the same spin degrees of freedom as $H_{\t{kin}}^{(\phi)}$.
-Nonetheless, for sufficiently weak SOC (small $s\phi$) the effect of $H_{\t{int}}^{(\phi)}$ may be approximated by $H_{\t{int}}^{(0)}=H_{\t{int}}$, which has previously been benchmarked for SU(2)-symmetric interactions \cite{he2019engineering, smale2019observation}.
-To ensure that $H_{\t{kin}}^{(\phi)}$ does not become trivial as $\phi\to0$, we can simultaneously take $J\to\infty$ ($J$ large compared to $U$) while keeping $J\phi$ constant.
+  \left. \H_{\t{kin}}^{(\phi)} \right|_{s\phi\ll1}
+  = 2J\phi \sum_q \sin\p{q} \s_{\z,q} + O\p{(s\phi)^2}.
+\end{align}
+For larger $\phi$, this Hamiltonian acquires terms with higher powers of $\s_{\z,q}$, up to $\s_{\z,q}^{n-1}$.
+
+In principle, making the gauge transformation $\c_{q\mu}^\dag\to \c_{q-\mu\phi,\mu}^\dag$ also transforms the interaction Hamiltonian as $\H_{\t{int}} \to \H_{\t{int}}^{(\phi)}$, where
+\begin{align}
+  \H_{\t{int}}^{(\phi)}
+  \equiv -\f{u}{2N} \sum_{p,q,\mu,\nu}
+  \c_{p-\mu\phi,\mu}^\dag \c_{p-\nu\phi,\nu}
+  \c_{q-\nu\phi,\nu}^\dag \c_{q-\mu\phi,\mu}.
+\end{align}
+However, $\H_{\t{int}}^{(\phi)}$ cannot be exactly reduced to a spin Hamiltonian that addresses the same spin degrees of freedom as $\H_{\t{kin}}^{(\phi)}$.
+Nonetheless, for sufficiently weak SOC (small $s\phi$) the effect of $\H_{\t{int}}^{(\phi)}$ may be approximated by $\H_{\t{int}}^{(0)}=\H_{\t{int}}$, which has previously been benchmarked for SU($2$)-symmetric interactions \cite{he2019engineering, smale2019observation}.
+To ensure that $\H_{\t{kin}}^{(\phi)}$ does not become trivial as $\phi\to0$, we can keep $J\phi/u$ constant (either by increasing $J/U$ or decreasing $N$).
 Altogether, the interacting spin Hamiltonian in the gauged frame becomes
 \begin{align}
-  H_\spin = -\f{U}{2N} \v S\c\v S + J\phi \sum_q \sin\p{q} s_{\z,q},
+  \H_\spin = -\f{u}{2N} \v\S\cdot\v\S + 2J\phi \sum_q \sin\p{q} \s_{\z,q},
   \label{eq:H_spin}
 \end{align}
-consisting of a spin-locking $\v S\c\v S$ term that energetically favors permutational symmetry, and an inhomogeneous magnetic field that causes inter-spin dephasing.
+consisting of a spin-locking $\v\S\cdot\v\S$ term that energetically favors permutational symmetry, and an inhomogeneous magnetic field that causes inter-spin dephasing.
 
@@
 An electronic ${^1}\t{S}_0\to {^3}\t{P}_0$ transition is thus accompanied by a nuclear spin transition $\mu_{\v v}\to\mu_{\v v}\pm1$, where $\mu_{\v v}$ is the projection of nuclear spin onto the axis $\v v$.
-Furthermore, a laser pointing along $\v v$ will generally imprint a phase $e^{-\i\kappa\v v\c\v\ell j}$ on lattice site $j$, where $\kappa$ is the laser's wavenumber (in units with lattice spacing $a=1$) and $\v\ell$ is a fixed unit vector parallel to the lattice.
+Furthermore, a laser pointing along $\v v$ will generally imprint a phase $e^{-\i\kappa\v v\cdot\v\ell j}$ on lattice site $j$, where $\kappa$ is the wavenumber of the laser (in units with lattice spacing $a=1$) and $\v\ell$ is a fixed unit vector parallel to the lattice.
 Altogether, the multi-laser drive Hamiltonian can be written in the form
 \begin{align}
-  H_{\t{drive}}^{\t{full}}
+  \H_{\t{drive}}^{\t{full}}
   = \sum_{j,\v v,\sigma} \Omega_{\v v\sigma}
-  \p{e^{-\i\kappa\v v\c\v\ell j} s_{\v v\sigma j}\otimes\op{\e}{\g}_j + \t{h.c.}}
-  + \Delta N_\e,
+  \p{e^{-\i\kappa\v v\cdot\v\ell j} \s_{\v v\sigma j}\otimes\op{\e}{\g}_j + \t{h.c.}}
+  + \Delta \hat N_\e,
   \label{eq:drive_full}
 \end{align}
-where $\sigma\in\set{+1,-1}$ indexes a polarization of light; $s_{\v v\sigma j}$ is a $n$-level spin-raising ($\sigma=+1$) or spin-lowering ($\sigma=-1$) operator along axis $\v v$ for atom $j$ (clarified below); $\ket\g_j$ and $\ket\e_j$ respectively denote the ground (${^1}\t{S}_0$) and excited (${^3}\t{P}_0$) electronic states of an atom on site $j$; and $N_\e=\1\otimes\sum_j\op\e_j$ is the number of excited atoms (with $\1$ the identity operator on all spin degrees of freedom).
-
-To define the spin-raising/lowering operators $s_{\v v\sigma j}$, we first define $s_{\x,j}$ similarly to $s_{\x,q}$ in Eq.~\eqref{eq:spin_x_op} with $q\leftrightarrow j$.
-The spin-raising operator $s_{+,j}$ is then twice the spin-raising part of $s_{\x,j}$, just as in the case of SU(2) where $s_\x=\frac12\op{\up}{\dn}+\t{h.c.}$, and the spin-lowering operator is $s_{-,j}\equiv s_{+,j}^\dag$.
-The operator $s_{\v v,+\,j}$ that raises spin along $\v v$ is acquired by a rotation of $s_{+,j}$ that maps the north pole to $\v v$:
-\begin{align}
-  s_{\v v,+} \equiv U_{\v v} s_+ U_{\v v}^\dag,
-  &&
-  U_{\v v} \equiv e^{-\i\alpha_{\v v}s_\z} e^{-\i\beta_{\v v}s_\y},
+where $\sigma\in\set{+1,-1}$ indexes a polarization of light; $\s_{\v v\sigma j}$ is a $n$-level spin-raising ($\sigma=+1$) or spin-lowering ($\sigma=-1$) operator along axis $\v v$ for atom $j$ (clarified below); $\ket\g_j$ and $\ket\e_j$ respectively denote the ground (${^1}\t{S}_0$) and excited (${^3}\t{P}_0$) electronic states of an atom on site $j$; and $\hat N_\e=\1\otimes\sum_j\op\e_j$ is the number of excited atoms (with $\1$ the identity operator on all spin degrees of freedom).
+
+To define the spin-raising/lowering operators $\s_{\v v\sigma j}$, we first define $\s_{\x,j}$ similarly to $\s_{\x,q}$ in Eq.~\eqref{eq:spin_x_op} with $q\leftrightarrow j$.
+The spin-raising operator $\s_{+,j}$ is then twice the spin-raising part of $\s_{\x,j}$, just as in the case of SU($2$) where $\s_\x=\frac12\op{\up}{\dn}+\t{h.c.}$, and the spin-lowering operator is $\s_{-,j}\equiv \s_{+,j}^\dag$.
+The operator $\s_{\v v,+\,j}$ that raises spin along $\v v$ is acquired by a rotation of $\s_{+,j}$ that maps the north pole to $\v v$:
+\begin{align}
+  \s_{\v v,+} \equiv \hat U_{\v v} \s_+ \hat U_{\v v}^\dag,
+  &&
+  \hat U_{\v v} \equiv e^{-\i\alpha_{\v v}\s_\z} e^{-\i\beta_{\v v}\s_\y},
   \label{eq:spin_rot}
 \end{align}
-where we have dropped the site index $j$ for brevity; $\alpha_{\v v}$ and $\beta_{\v v}$ are the azimuthal and polar angles of $\v v$; and $s_\y\equiv-\frac\i2s_++\t{h.c.}$.
-Note that the operators $(s_\x,s_\y,s_\z)$ form an $\su(2)$ algebra, so given a particular axis $\v v$ one can expand $s_{\v v,+}=\sum_m c_{\v vm} s_m$ with coefficients $c_{\v vm}$ that are independent of the dimension $n$.
-In particular, one is always free to set $n=2$ and expand Eq.~\eqref{eq:spin_rot} with spin-$\frac12$ operators (or Pauli matrices) to find the coefficients $c_{\v vm}$.
+where we have dropped the site index $j$ for brevity; $\alpha_{\v v}$ and $\beta_{\v v}$ are the azimuthal and polar angles of $\v v$; and $\s_\y\equiv-\frac\i2\s_++\t{h.c.}$.
+Note that the operators $(\s_\x,\s_\y,\s_\z)$ form an $\su(2)$ algebra, so given a particular axis $\v v$ one can expand $\s_{\v v,+}=\sum_m f_{\v vm} \s_m$ with coefficients $f_{\v vm}$ that are independent of the dimension $n$.
+In particular, one is always free to set $n=2$ and expand Eq.~\eqref{eq:spin_rot} with spin-$\frac12$ operators (or Pauli matrices) to find the coefficients $f_{\v vm}$.
 
@@
 
-In order to make the drive Hamiltonian in Eq.~\eqref{eq:drive_full} more tractable, we orient the lattice along the $z$ axis, $\v\ell=(0,0,1)$, and consider a three-laser drive with a geometry sketched in Figure \ref{fig:3LD}.
-In this setup, two counter-propagating right-circularly polarized lasers point along $\v v_\pm=\pm(0,-\sin\theta,\cos\theta)$, at an angle $\theta$ to the lattice axis, and a third linearly-polarized laser points along $\v v_0=(1,0,0)$.
-Choosing $\v v_+$ as the spin quantization axis, we can define new drive amplitudes $\Omega_0\equiv-(\Omega_{\v v_0,+}+\Omega_{\v v_0,-})$ and $\Omega_\pm\equiv\pm\Omega_{\v v_\pm,+}$, as well as a SOC angle $\phi\equiv\kappa\v v_+\c\v\ell=\kappa\cos\theta$, in terms of which the drive Hamiltonian becomes
-\begin{align}
-  H_{\t{3LD}}^{\t{full}}
+We now consider a specific three-laser driving scheme with a geometry sketched in Figure \ref{fig:3LD}.
+Here the lattice lies in the y-z plane at an angle $\theta$ to the z axis, oriented along $\v\ell=(0,\sin\theta,\cos\theta)$; two counter-propagating right-circularly polarized lasers point in opposite directions along the z axis, $\v v_\pm=(0,0,\pm1)$; and a third linearly-polarized laser points along the x axis, $\v v_0=(1,0,0)$.
+Choosing $\v v_+$ (i.e.~the z axis) as the spin quantization axis, we can define new drive amplitudes $\Omega_0\equiv-(\Omega_{\v v_0,+}+\Omega_{\v v_0,-})$ and $\Omega_\pm\equiv\pm\Omega_{\v v_\pm,+}$, as well as a SOC angle $\phi\equiv\kappa\v v_+\cdot\v\ell=\kappa\cos\theta$, in terms of which the drive Hamiltonian becomes
+\begin{align}
+  \H_{\t{3LD}}^{\t{full}}
   = \sum_{j,m} \Omega_m
-  \p{e^{-\i m\phi j} s_{mj} \otimes\op{\e}{\g}_j + \t{h.c.}}
-  + \Delta N_\e,
-\end{align}
-where $m\in\set{+1,0,-1}$ indexes the laser pointing along $\v v_m$; and $s_{0,j}\equiv s_{\z,j}$ for shorthand.
+  \p{e^{-\i m\phi j} \s_{mj} \otimes\op{\e}{\g}_j + \t{h.c.}}
+  + \Delta \hat N_\e,
+\end{align}
+where $m\in\set{+1,0,-1}$ indexes the laser pointing along $\v v_m$; and $\s_{0,j}\equiv \s_{\z,j}$ for shorthand.
 In the far-detuned limit $\abs{\Delta}\gg\abs{\Omega_m}$, a second-order perturbative treatment of electronic excitations ($\ket\e$) yields an effective drive Hamiltonian that only addresses ground-state nuclear spins.
-After additionally making the gauge transformation $c_{j\mu}^\dag\to e^{\i\phi\mu j} c_{j\mu}^\dag$ (equivalently $s_{mj}\to e^{\i m\phi j}s_{mj}$), the drive Hamiltonian then becomes
-\begin{align}
-  H_{\t{3LD}} = \sum_j H_{\t{3LD},j}^{\t{single}},
+After additionally making the gauge transformation $\c_{j\mu}^\dag\to e^{\i\phi\mu j} \c_{j\mu}^\dag$ (equivalently $\s_{mj}\to e^{\i m\phi j}\s_{mj}$), the drive Hamiltonian then becomes
+\begin{align}
+  \H_{\t{3LD}} = \sum_j \H_{\t{3LD},j}^{\t{single}},
   \label{eq:drive_all}
 \end{align}
-where $H_{\t{3LD},j}^{\t{single}}$ denotes the action of $H_{\t{3LD}}^{\t{single}}$ on spin $j$:
+where $\H_{\t{3LD},j}^{\t{single}}$ denotes the action of $\H_{\t{3LD}}^{\t{single}}$ on spin $j$:
 \begin{multline}
-  H_{\t{3LD}}^{\t{single}}
-  = \tilde\Omega_+ \tilde\Omega_- s_\z
-  + \tilde\Omega_0 \tilde\Omega_- s_\x
-  + \tilde\Omega_0 \tilde\Omega_+ \p{s_\z s_\x  + s_\x s_\z} \\
-  - \tilde\Omega_0^2 s_\z^2 - \tilde\Omega_+^2 s_\x^2
-  - \tilde\Omega_-^2 s_\y^2,
+  \H_{\t{3LD}}^{\t{single}}
+  = \tilde\Omega_+ \tilde\Omega_- \s_\z
+  + \tilde\Omega_0 \tilde\Omega_- \s_\x
+  + \tilde\Omega_0 \tilde\Omega_+ \p{\s_\z \s_\x  + \s_\x \s_\z} \\
+  - \tilde\Omega_0^2 \s_\z^2 - \tilde\Omega_+^2 \s_\x^2
+  - \tilde\Omega_-^2 \s_\y^2,
   \label{eq:drive_single}
@@
 \end{align}
-where we have made the simplifying assumption that all drive amplitudes are real to arrive at the form of $H_{\t{3LD}}^{\t{single}}$ in Eq.~\eqref{eq:drive_single}.
+where we have made the simplifying assumption that all drive amplitudes are real to arrive at the form of $\H_{\t{3LD}}^{\t{single}}$ in Eq.~\eqref{eq:drive_single}.
 We relax the assumption of real drive amplitudes in Appendix \ref{sec:full_drive}.
@@
 Drive Hamiltonians (left column) that can be implemented with different amplitude-matching conditions (right column), some of which are specified by an arbitrary sign $\sigma\in\set{+1,-1}$.
-The drives shown here are equal to that Eq.~\eqref{eq:drive_single} up to a possible energy shift of $s_\x^2+s_\y^2+s_\z^2=s(s+1)$, and come in mutually commuting pairs: a drive with $\Omega_m=1$ and $\Omega_n=0$ for both $n\ne m$ commutes with the drive in which $\Omega_m=0$ and both $\abs*{\Omega_n}=1$.
+The drives shown here are equal to that Eq.~\eqref{eq:drive_single} up to a possible energy shift of $\s_\x^2+\s_\y^2+\s_\z^2=s(s+1)$, and come in mutually commuting pairs: a drive with $\Omega_m=1$ and $\Omega_n=0$ for both $n\ne m$ commutes with the drive in which $\Omega_m=0$ and both $\abs*{\Omega_n}=1$.
 }
@@
 \begin{tabular}{c|c}
-  $H_{\t{drive}}^{\t{single}}$
+  $\H_{\t{drive}}^{\t{single}}$
   & $(\tilde\Omega_0,\tilde\Omega_+,\tilde\Omega_-)$
   \\ \hline\hline
-  $-s_\z^2$ & $\p{1,0,0}$
+  $-\s_\z^2$ & $\p{1,0,0}$
   \\ \hline
-   $-s_\x^2$ & $\p{0,1,0}$
+   $-\s_\x^2$ & $\p{0,1,0}$
   \\ \hline
-  $-s_\y^2$ & $\p{0,0,1}$
+  $-\s_\y^2$ & $\p{0,0,1}$
   \\ \hline
-  $\sigma s_\z + s_\z^2$ & $\p{0,1,\sigma}$
+  $\sigma \s_\z + \s_\z^2$ & $\p{0,1,\sigma}$
   \\ \hline
-  $\sigma s_\x + s_\x^2$ & $\p{1,0,\sigma}$
+  $\sigma \s_\x + \s_\x^2$ & $\p{1,0,\sigma}$
   \\ \hline
-  $\sigma\p{s_\z s_\x+s_\x s_\z} + s_\y^2$ & $\p{1,\sigma,0}$
+  $\sigma\p{\s_\z \s_\x+\s_\x \s_\z} + \s_\y^2$ & $\p{1,\sigma,0}$
   \\ \hline
-  $\pm s_\z \pm \sigma s_\x + \sigma \p{s_\z s_\x + s_\x s_\z}$
+  $\pm \s_\z \pm \sigma \s_\x + \sigma \p{\s_\z \s_\x + \s_\x \s_\z}$
   & $\p{1,\sigma,\pm\sigma}$
@@
 There are three important observations to make about these results.
-First, the fact that $H_{\t{3LD}}$ acts identically on all spins means we can freely replace the site index $j$ with a momentum index $q$ (as can be verified by substituting $c_{j\mu}=\frac1{\sqrt{L}}\sum_k e^{-\i q\c j} c_{q\mu}$), which is important to ensure that this drive addresses the same spin degrees of freedom as the spin Hamiltonian $H_\spin$ in Eq.~\eqref{eq:H_spin}.
+First, the fact that $\H_{\t{3LD}}$ acts identically on all spins means we can freely replace the site index $j$ with a momentum index $q$ (as can be verified by substituting $\c_{j\mu}=\frac1{\sqrt{L}}\sum_k e^{-\i q\cdot j} \c_{q\mu}$), which is important to ensure that this drive addresses the same spin degrees of freedom as the spin Hamiltonian $\H_\spin$ in Eq.~\eqref{eq:H_spin}.
 Second, each of $\tilde\Omega_0,\tilde\Omega_+,\tilde\Omega_-$ can be tuned independently by changing the amplitudes of the driving lasers; some particular Hamiltonians for specific values of these amplitudes are shown in Table \ref{tab:drives}.
-Third, due to the appearance of mutually commuting pairs of Hamiltonians in Table \ref{tab:drives}, specifically $-s_\alpha^2$ and $\pm s_\alpha+s_\alpha^2$ for $\alpha\in\set{\z,\x}$, the three-laser drive admits pulse sequences that exactly implement arbitrary SU(2) (spatial) rotations of the form $e^{-\i\chi\vec n\c\vec s}$, where $\chi$ is a rotation angle, $\vec n$ is a rotation axis, and $\vec s\equiv(s_\x,s_\y,s_\z)$.
-The capability to perform arbitrary spatial rotations, together with the capability to measure the number of atoms with spin projection $\mu$ onto a fixed quantization axis, $\bk{S_{\mu\mu}}$ (where $S_{\mu\nu}=\sum_js_{\mu\nu j}$), implies the capability to reconstruct all components of the mean collective spin matrix $\bk{\v S}=\sum_{\mu\nu}\bk{S_{\mu\nu}}\op{\mu}{\nu}$ via spin qudit tomography \cite{newton1968measurability, perlin2020qudit}.
+Third, due to the appearance of mutually commuting pairs of Hamiltonians in Table \ref{tab:drives}, specifically $-\s_\alpha^2$ and $\pm \s_\alpha+\s_\alpha^2$ for $\alpha\in\set{\z,\x}$, the three-laser drive admits pulse sequences that exactly implement arbitrary SU($2$) (spatial) rotations of the form $e^{-\i\chi\vec n\cdot\vec s}$, where $\chi$ is a rotation angle, $\vec n$ is a rotation axis, and $\vec s\equiv(\s_\x,\s_\y,\s_\z)$.
+The capability to perform arbitrary spatial rotations, together with the capability to measure the number of atoms with spin projection $\mu$ onto a fixed quantization axis, $\bk{\S_{\mu\mu}}$ (where $\S_{\mu\nu}=\sum_j\s_{\mu\nu j}$), implies the capability to reconstruct all components of the mean collective spin matrix $\bk{\v\S}=\sum_{\mu\nu}\bk{\S_{\mu\nu}}\op{\mu}{\nu}$ via spin qudit tomography \cite{newton1968measurability, perlin2020qudit}.
 Moreover, we expect that advanced quantum optimal control techniques (similar to those of Refs.~\cite{anderson2015accurate, lucarelli2018quantum}) can be used to implement arbitrary SU($n$) rotations by designing suitable time-dependent drive amplitudes $\Omega_m$.
@@
 Finally, we comment on the preparation of initial states.
-Initial states are nominally prepared in the ``lab frame'', and must be transformed according to the gauge transformation $c_{q\mu}^\dag\to c_{q-\mu\phi,\mu}^\dag$ prior to evolution under the ``gauged frame'' spin Hamiltonian $H_\spin$ in Eq.~\eqref{eq:H_spin} or the three-laser drive $H_{\t{3LD}}$ in Eq.~\eqref{eq:drive_all}.
-We assume the capability to prepare an initial state in which all spins are maximally polarized along $\v v_+$, i.e.~$\ket{\v v_+}^{\otimes N}=\ket{s}^{\otimes N}$, which is unaffected by the gauge transformation (up to a global phase).
+Initial states are nominally prepared in the ``lab frame'', and must be transformed according to the gauge transformation $\c_{q\mu}^\dag\to \c_{q-\mu\phi,\mu}^\dag$ prior to evolution under the ``gauged frame'' spin Hamiltonian $\H_\spin$ in Eq.~\eqref{eq:H_spin} or the three-laser drive $\H_{\t{3LD}}$ in Eq.~\eqref{eq:drive_all}.
+We assume the capability to prepare an initial state in which all spins are maximally polarized along the z axis, i.e.~$\ket{\z}^{\otimes N}=\ket{s}^{\otimes N}$, which is unaffected by the gauge transformation (up to a global phase).
 The three-laser then allows us to rotate this state into one that is polarized along any spatial axis (in the gauged frame).
@@
 \begin{align}
-  e^{-\i\frac{\pi}{2}\p{s_\y + s_\y^2}} \ket{s}
+  e^{-\i\frac{\pi}{2}\p{\s_\y + \s_\y^2}} \ket{s}
   \propto \ket{s} + \ket{-s}.
 \end{align}
-This state has a vanishing mean spin vector, $\bk{s_\x}=\bk{s_\y}=\bk{s_\z}=0$, but variances $\bk{s_\x^2}=\bk{s_\y^2}=s/2$ and $\bk{s_\z^2}=s^2$.
+This state has a vanishing mean spin vector, $\bk{\s_\x}=\bk{\s_\y}=\bk{\s_\z}=0$, but variances $\bk{\s_\x^2}=\bk{\s_\y^2}=s/2$ and $\bk{\s_\z^2}=s^2$.
 
@@
 
-We now return to the spin Hamiltonian $H_\spin$ in Eq.~\eqref{eq:H_spin}, which is valid in the weak-SOC limit $s\phi\ll1$, and henceforth work exclusively in the ``gauged frame'' of $H_\spin$ and the three-laser drive $H_{\t{3LD}}$ in Eq.~\eqref{eq:drive_all}.
+We now return to the spin Hamiltonian $\H_\spin$ in Eq.~\eqref{eq:H_spin}, which is valid in the weak-SOC limit $s\phi\ll1$, and henceforth work exclusively in the ``gauged frame'' of $\H_\spin$ and the three-laser drive $\H_{\t{3LD}}$ in Eq.~\eqref{eq:drive_all}.
 At the mean-field (MF) level, up to constant energy shifts the undriven spin Hamiltonian becomes
 \begin{align}
-  H_\MF = \sum_q\sp{-U\bk{\bar{\v s}}\c\v s_q
-    + J\phi \sum_q \sin\p{q} s_{\z,q}}.
+  \H_\MF = u \sum_q\sp{-\bk{\bar{\v s}}\cdot\v\s_q
+    + h \sum_q \sin\p{q} \s_{\z,q}}.
   \label{eq:H_MF}
 \end{align}
-where $\bar{\v s}\equiv\frac1N\sum_q\v s_q$ is the average spin matrix.
+where $\bar{\v s}\equiv\frac1N\sum_q\v\s_q$ is the average spin matrix, and $h\equiv 2J\phi/u$ is a dimensionless strength of the inhomogeneous magnetic field.
 We assume that all momenta $q\in\ZZ_N\times 2\pi/N$ are occupied.
-Fixing the atom number $N$, the spin Hamiltonian has one free parameter: the reduced field $J\phi/U$ that determines the relative strength of the single-particle and interaction terms.
-One should therefore expect distinct dynamical behaviors when $J\phi/U\ll1$, in which case strong spin-locking interactions should give rise to a long-range ordered phase, as opposed to $J\phi/U\gg1$, in which case long-range order should be destroyed by the strong inhomogeneous magnetic field.
+Fixing the atom number $N$, the spin Hamiltonian has one free parameter, $h$, which determines the relative strength of the single-particle and interaction terms.
+One should therefore expect distinct dynamical behaviors when $h\ll1$, in which case strong spin-locking interactions should give rise to a long-range ordered phase, as opposed to $h\gg1$, in which case long-range order should be destroyed by the strong inhomogeneous magnetic field.
 
@@
   &&
-  \vec\sigma \equiv \f1{Ns} \times \p{S_\x,S_\y,S_\z},
-\end{align}
-where $S_\alpha \equiv \sum_q s_{\alpha,q}$, and the time-averaged (dimensionless) interaction energy
-\begin{align}
-  \bbk{\ss}_\MF = \f1{N^2} \times \bbk{\v S\c\v S}_\MF.
+  \vec\sigma \equiv \f1{Ns} \times \vec S,
+\end{align}
+where $\vec S\equiv(\S_\x,\S_\y,\S_\z)$ with $\S_\alpha \equiv \sum_q \s_{\alpha,q}$, and the time-averaged (dimensionless) interaction energy
+\begin{align}
+  \bbk{\ss}_\MF = \f1{N^2} \times \bbk{\v\S\cdot\v\S}_\MF.
 \end{align}
@@
 Time-averaged magnetization $\sigma_\MF$ and interaction energy $\bbk{\ss}_\MF$ for different spin dimensions $n$ (indicated in the legend) as determined by mean-field simulations of $N=100$ spins initially in the $x$-polarized state $\ket\X$ for a time $T=10^5/U$.
-Insets show same data after rescaling $J\phi/U\to J\phi/U \times (n/2)^{1/3}$, and transforming vertical axes according to Eq.~\eqref{eq:rescale}.
+Insets show same data after rescaling $h\to h \times (n/2)^{1/3}$, and transforming vertical axes according to Eq.~\eqref{eq:rescale}.
 }
@@
 
-Figure \ref{fig:mean_mag_int_X} shows the time-averages of the magnetization $\sigma_\MF$ and interaction energy $\bbk{\ss}_\MF$ as computed by mean-field simulations of $N=100$ spins initially in the $x$-polarized state $\X\equiv\ket\x^{\otimes N}$, where
-\begin{align}
-  \ket\x \equiv e^{-\i\frac{\pi}{2}s_\y} \ket{s}
+Figure \ref{fig:mean_mag_int_X} shows the time-averages of the magnetization $\sigma_\MF$ and interaction energy $\bbk{\ss}_\MF$ as computed by mean-field simulations of $N=100$ spins initially in the $x$-polarized state $\ket\X\equiv\ket\x^{\otimes N}$, where
+\begin{align}
+  \ket\x \equiv e^{-\i\frac{\pi}{2}\s_\y} \ket{s}
   = \f1{2^s} \sum_\mu { 2s \choose s+\mu }^{1/2} \ket{\mu}.
@@
 Here ${ m \choose k }$ is a binomial coefficient.
-As expected, the spin model exhibits a mean-field dynamical phase transition between an ordered phase at small $J\phi/U$ and a disordered phase at large $J\phi/U$.
-The ordered phase has a non-zero magnetization $\sigma_\MF$ and an interaction energy $\bbk{\ss}_\MF$ that asymptotically approach their maximal values as $J\phi/U\to0$.
+As expected, the spin model exhibits a mean-field dynamical phase transition between an ordered phase at small $h$ and a disordered phase at large $h$.
+The ordered phase has a non-zero magnetization $\sigma_\MF$ and an interaction energy $\bbk{\ss}_\MF$ that asymptotically approach their maximal values as $h\to0$.
 The disordered phase has no (time-averaged) magnetization, $\sigma_\MF=0$, but the interaction energy $\bbk{\ss}_\MF$ nonetheless indicates persistent nontrivial inter-spin correlations when $n>2$.
-These nontrivial correlations vanish as $J\phi/U\to\infty$, in which case $\bbk{\ss}_\MF$ approaches the minimal value allowed by conservation laws (clarified below).
-By minimizing the reduced field $J\phi/U$ for which $\sigma_\MF=0$, we numerically find that the transition between ordered and disordered phases occurs at a critical field $\p{J\phi/U}_\crit=\p{n/2}^{-\alpha}$ with $\alpha\approx1/3$ (see Figure \ref{fig:crit_fields_X}).
-When $n=2$, this transition is consistent with the predictions of a Lax vector analysis \cite{yuzbashyan2005nonequilibrium, yuzbashyan2006dynamical, yuzbashyan2006relaxation, yuzbashyan2015quantum, smale2019observation} that exploits integrability of $H_\spin$ to determine long-time behavior.
-However, a straightforward application of the same analysis fails to predict the mean-field transition for $n>2$, so additional theoretical tools are necessary to understand this transition.
+These nontrivial correlations vanish as $h\to\infty$, in which case $\bbk{\ss}_\MF$ approaches the minimal value allowed by conservation laws (clarified below).
+By minimizing the reduced field $h$ for which $\sigma_\MF=0$, we numerically find that the transition between ordered and disordered phases occurs at a critical field $h_\crit=\p{n/2}^{-\alpha}$ with $\alpha\approx1/3$ (see Figure \ref{fig:crit_fields_X}).
+When $n=2$, this transition is consistent with the predictions of a Lax vector analysis \cite{yuzbashyan2005nonequilibrium, yuzbashyan2006dynamical, yuzbashyan2006relaxation, yuzbashyan2015quantum, smale2019observation} that exploits integrability of $\H_\spin$ to determine long-time behavior.
+However, additional theoretical tools are necessary to understand this transition when $n>2$.
 We elaborate on this point in Appendix \ref{sec:lax}.
@@
 \caption{
-The critical value of $\p{J\phi/U}_\crit$ as determined by mean-field simulations of $N=100$ spins initially in the $x$-polarized state $\ket\X$.
-A single-parameter fit to $\p{J\phi/U}_\crit=\p{n/2}^{-\alpha}$ finds $\alpha=0.333(5)$, and $\alpha=1/3$ is consistent with all mean-field results to within an uncertainty determined by the resolution of $J\phi/U$ in mean-field simulations.
+The critical value of $h_\crit$ as determined by mean-field simulations of $N=100$ spins initially in the $x$-polarized state $\ket\X$.
+A single-parameter fit to $h_\crit=\p{n/2}^{-\alpha}$ finds $\alpha=0.333(5)$, and $\alpha=1/3$ is consistent with all mean-field results to within an uncertainty determined by the resolution of $h$ in mean-field simulations.
 }
@@
 
-As shown in insets of Figure \ref{fig:mean_mag_int_X}, mean-field results for different spin dimensions $n$ collapse onto each other when normalizing the field $J\phi/U$ to its critical value, $\p{J\phi/U}\to\p{J\phi/U}\times\p{n/2}^{1/3}$, and rescaling
+As shown in insets of Figure \ref{fig:mean_mag_int_X}, mean-field results for different spin dimensions $n$ collapse onto each other when normalizing the field $h$ to its critical value, $h\to h\times\p{n/2}^{1/3}$, and rescaling
 \begin{align}
@@
 \end{align}
-The rescaling of magnetization and interaction energy can be understood by considering their limiting behavior as $J\phi/U\to\infty$ or $J\phi/U\to0$.
-
-In the strong-field limit $J\phi/U\to\infty$, we can ignore interactions and treat spins as though they simply precess at different rates.
-The time-averaged transverse magnetization $\sigma_\MF$ then trivially vanishes as $J\phi/U\to\infty$.
-The interaction energy $\bk{\ss}_\MF = \bk{\bar{\v s}}_\MF \c \bk{\bar{\v s}}_\MF$, meanwhile, has contributions from
+The rescaling of magnetization and interaction energy can be understood by considering their limiting behavior as $h\to\infty$ or $h\to0$.
+
+In the strong-field limit $h\to\infty$, we can ignore interactions and treat spins as though they simply precess at different rates.
+The time-averaged transverse magnetization $\sigma_\MF$ then trivially vanishes as $h\to\infty$.
+The interaction energy $\bk{\ss}_\MF = \bk{\bar{\v s}}_\MF \cdot \bk{\bar{\v s}}_\MF$, meanwhile, has contributions from:
 \begin{enumerate*}
@@
 \begin{align}
-  \lim_{J\phi/U\to\infty} \bbk{\ss}_\MF
+  \lim_{h\to\infty} \bbk{\ss}_\MF
   = \Tr\sp{\p{\diag\op{\x}}^2}
@@
 
-In the weak-field limit $J\phi/U\to0$, the spin-locking $\v S\c\v S$ interactions of the spin Hamiltonian $H_\spin$ energetically restrict dynamics to the permutationally symmetric (PS) manifold.
+In the weak-field limit $h\to0$, the spin-locking $\v\S\cdot\v\S$ interactions of the Hamiltonian $\H_\spin$ energetically restrict dynamics to the permutationally symmetric (PS) manifold.
 At second order in perturbation theory, the effective spin Hamiltonian within the PS manifold becomes (see Appendix \ref{sec:pert_theory})
 \begin{align}
-  H_\spin^\eff
-  = \f{\p{J\phi}^2}{2\p{N-1}U}
-  \times \sp{S_\z^2 - N\sum_q s_{\z,q}^2},
+  \H_\spin^\eff
+  = \f{h^2 u}{2(N-1)} \times \sp{\S_\z^2 - N\sum_q \s_{\z,q}^2},
   \label{eq:H_spin_eff}
 \end{align}
-whose mean-field version is
-\begin{align}
-  H_\MF^\eff = -\f{\p{J\phi}^2}{2U} \sum_q s_{\z,q}^2,
-\end{align}
-where we have used the fact that the axial magnetizations $\bk{s_{\z,q}}=\frac1N\bk{S_\z}$ within the PS manifold, and the initial value of $\bk{S_\z}=0$ is conserved by $H_\spin$.
-The weak-field effective Hamiltonian preserves permutational symmetry, so $\bbk{\ss}_\MF\to1$ as $J\phi/U\to0$.
-Moreover, both the spin Hamiltonian $H_\spin$ and the initial state $\ket\X$ are invariant (up to global phase) under the action of $\Pi_\z R_\x^\pi$, where $R_\x^\theta \equiv e^{-\i\theta S_\x}$ is a rotation about the $x$ axis by $\theta$ and $\Pi_\z\equiv\sum_{q>0}\v s_q\c\v s_{-q}$ is a permutation of spins $q\leftrightarrow-q$\footnote{$\Pi_\z$ can also be thought of a reflection operator that inverts momentum along the $z$ axis.}, which is to say that
-\begin{align}
-  e^{-\i t H_\spin} \ket\X
-  &\simeq \p{\Pi_\z R_\x^\pi e^{-\i t H_\spin} {R_\x^\pi}^\dag \Pi_\z^\dag} \Pi_\z R_\x^\pi \ket\X
+which in the mean-field approximation becomes
+\begin{align}
+  \H_\MF^\eff = -\frac12 h^2 u \sum_q \s_{\z,q}^2,
+\end{align}
+where we have used the fact that the axial magnetizations $\bk{\s_{\z,q}}=\frac1N\bk{\S_\z}$ within the PS manifold, and the initial value of $\bk{\S_\z}=0$ is conserved by $\H_\spin$.
+The weak-field effective Hamiltonian preserves permutational symmetry, so $\bbk{\ss}_\MF\to1$ as $h\to0$.
+Moreover, both the spin Hamiltonian $\H_\spin$ and the initial state $\ket\X$ are invariant (up to global phase) under the action of $\hat\Pi_\z \R_\x^\pi$, where $\R_\x^\theta \equiv e^{-\i\theta \S_\x}$ is a rotation about the x axis by $\theta$ and $\hat\Pi_\z\equiv\sum_{q>0}\v\s_q\cdot\v\s_{-q}$ is a permutation of spins $q\leftrightarrow-q$\footnote{$\hat\Pi_\z$ can also be thought of a reflection operator that inverts momentum along the z axis.}, which is to say that
+\begin{align}
+  e^{-\i t \H_\spin} \ket\X
+  &\simeq \p{\hat\Pi_\z \R_\x^\pi e^{-\i t \H_\spin} \R_\x^\pi{}^\dag \hat\Pi_\z^\dag} \hat\Pi_\z \R_\x^\pi \ket\X
   \label{eq:sym_x_start} \\
-  &\simeq \Pi_\z R_\x^\pi e^{-\i t H_\spin} \ket\X,
+  &\simeq \hat\Pi_\z \R_\x^\pi e^{-\i t \H_\spin} \ket\X,
 \end{align}
 where $\simeq$ denotes equality up to a global phase.
-The permutation $\Pi_\z$ commutes with $S_\y$, while the rotation $R_\x^\pi$ takes $S_\y\to-S_\y$, so
-\begin{align}
-  \bk{S_\y}
-  = \bk{{R_\x^\pi}^\dag \Pi_\z^\dag S_\y \Pi_\z R_\x^\pi}
-  = -\bk{S_\y} = 0
+The permutation $\hat\Pi_\z$ commutes with $\S_\y$, while the rotation $\R_\x^\pi$ takes $\S_\y\to-\S_\y$, so
+\begin{align}
+  \bk{\S_\y}
+  = \bk{\R_\x^\pi{}^\dag \hat\Pi_\z^\dag \S_\y \hat\Pi_\z \R_\x^\pi}
+  = -\bk{\S_\y} = 0
   \label{eq:sym_x_end}
@@
 at all times.
-In turn, conservation of $\bk{S_\y}=\bk{S_\z}=0$ and permutational symmetry together imply that the magnetization $\sigma_\MF$ is determined by the time-average of $\sigma_\x\equiv s_\x/s$ for a single spin:
-\begin{align}
-  \lim_{J\phi/U\to0} \sigma_\MF
+In turn, conservation of $\bk{\S_\y}=\bk{\S_\z}=0$ and permutational symmetry together imply that the magnetization $\sigma_\MF$ is determined by the time-average of $\hat\sigma_\x\equiv\s_\x/s$ for a single spin:
+\begin{align}
+  \lim_{h\to0} \sigma_\MF
   = \abs{\lim_{T\to\infty} \f1T \int_0^T \dd\tau
-  \bk{\x|\sigma_\x\p{\tau}|\x}},
+  \bk{\x|\hat\sigma_\x\p{\tau}|\x}},
 \end{align}
@@
 \begin{align}
-  \sigma_\x\p{\tau} = e^{\i\tau s_\z^2} \sigma_\x e^{-\i\tau s_\z^2}.
-\end{align}
-By mapping the Hilbert space of an $n$-level spin onto the PS (Dicke) manifold of $n-1$ spin-$\frac12$ particles, we can adapt exact analytical results for the Ising model \cite{foss-feig2013nonequilibrium}\footnote{See Appendix K of Ref.~\cite{perlin2020shorttime} for a simpler adaptation of the analytics in Ref.~\cite{foss-feig2013nonequilibrium} to the one-axis twisting model $H_{\t{OAT}}=\chi s_\z^2$.} to find that
-\begin{align}
-  \bk{\x|\sigma_\x\p{\tau}|\x} = \p{\cos\tau}^{n-2},
+  \hat\sigma_\x\p{\tau} = e^{\i\tau \s_\z^2} \hat\sigma_\x e^{-\i\tau \s_\z^2}.
+\end{align}
+By mapping the Hilbert space of an $n$-level spin onto the PS (Dicke) manifold of $n-1$ spin-$\frac12$ particles, we can adapt exact analytical results for the Ising model \cite{foss-feig2013nonequilibrium}\footnote{See Appendix K of Ref.~\cite{perlin2020shorttime} for a simpler adaptation of the analytics in Ref.~\cite{foss-feig2013nonequilibrium} to the one-axis twisting model $\H_{\t{OAT}}=\chi \s_\z^2$.} to find that
+\begin{align}
+  \bk{\x|\hat\sigma_\x\p{\tau}|\x} = \p{\cos\tau}^{n-2},
 \end{align}
@@
 \begin{align}
-  \lim_{J\phi/U\to0} \sigma_\MF
+  \lim_{h\to0} \sigma_\MF
   = \f1{2\pi} \int_0^{2\pi} \dd\tau \p{\cos\tau}^{n-2}
@@
 \end{align}
-When going beyond mean-field theory, inter-spin correlations generated by $S_\z^2$ in Eq.~\eqref{eq:H_spin_eff} will cause $\bk{S_\x}$ to decay on time scales that diverge as $N\to\infty$, so the mean-field magnetization $\sigma_\MF$ should be interpreted as an ``intermediate-time-average'' of $\vec\sigma$.
+When going beyond mean-field theory, inter-spin correlations generated by $\S_\z^2$ in Eq.~\eqref{eq:H_spin_eff} will cause $\bk{\S_\x}$ to decay on time scales that diverge as $N\to\infty$, so the mean-field magnetization $\sigma_\MF$ should be interpreted as an ``intermediate-time-average'' of $\vec\sigma$.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Multi-cat initial states}
+\subsection{Kitten initial states}
 
@@
 
-We now consider the same setup as above, but with the initial ``multi-cat'' states $\ket\XX\equiv\ket\xx^{\otimes N}$ and $\ket\XXI\equiv\ket\xxi^{\otimes N}$, where
+We now consider the same setup as above, but with the initial ``kitten'' states $\ket\XX\equiv\ket\xx^{\otimes N}$ and $\ket\XXI\equiv\ket\xxi^{\otimes N}$, where
 \begin{align}
@@
 \end{align}
-and $\ket{-\x}$ is a state polarized along $-x$, defined similarly to $\ket\x$ in Eq.~\eqref{eq:state_x}:
-\begin{align}
-  \ket{-\x} \equiv e^{-\i\frac{\pi}{2}s_\y}\ket{-s}
+and $\ket{-\x}$ is a state polarized along $-$x, defined similarly to $\ket\x$ in Eq.~\eqref{eq:state_x}:
+\begin{align}
+  \ket{-\x} \equiv e^{-\i\frac{\pi}{2}\s_\y}\ket{-s}
   = \f1{2^s} \sum_\mu \p{-1}^{s+\mu}
@@
 Similarly to Figure \ref{fig:mean_mag_int_X}, Figures \ref{fig:mean_mag_int_XX} and \ref{fig:mean_mag_int_XXI} show the time-averaged magnetization $\sigma_\MF$ and interaction energy $\bbk{\ss}_\MF$ throughout mean-field dynamics of the initial states $\ket\XX$ and $\ket\XXI$.
-These figures exclude the trivial case of spin dimension $n=2$, for which $\ket\xx=\ket{-s}$ is an eigenstate of $H_\spin$ and $\ket\xxi=e^{-\i\frac{\pi}{2}s_\z}\ket\x\equiv\ket\y$ is spin-polarized along $y$.
+These figures exclude the trivial case of spin dimension $n=2$, for which $\ket\xx=\ket{-s}$ is an eigenstate of $\H_\spin$ and $\ket\xxi=e^{-\i\frac{\pi}{2}\s_\z}\ket\x\equiv\ket\y$ is spin-polarized along $y$.
 The first and perhaps most interesting observation to make about Figures \ref{fig:mean_mag_int_XX} and \ref{fig:mean_mag_int_XXI} is that they are different, signifying the importance of intra-spin coherences for the dynamical behavior of multilevel spin models.
 
-Unlike Figure \ref{fig:mean_mag_int_X} (for $\ket\X$), Figure \ref{fig:mean_mag_int_XX} (for $\ket\XX$) exhibits no sharp transition between distinct dynamical phases: the time-averaged magnetization $\sigma_\MF=0$ for all values of the field $J\phi/U$, and the interaction energy $\bbk{\ss}_\MF$ smoothly crosses over from a maximal value of 1 to a minimal value of $2\gamma\p{n}$.
-The minimal value of $\bbk{\ss}_\MF$ approached as $J\phi/U\to\infty$ can be explained with arguments identical to those in the paragraph containing Eq.~\eqref{eq:lim_ss}, which now imply that
-\begin{align}
-  \lim_{J\phi/U\to\infty} \bbk{\ss}_\MF
+Unlike Figure \ref{fig:mean_mag_int_X} (for $\ket\X$), Figure \ref{fig:mean_mag_int_XX} (for $\ket\XX$) exhibits no sharp transition between distinct dynamical phases: the time-averaged magnetization $\sigma_\MF=0$ for all values of the field $h$, and the interaction energy $\bbk{\ss}_\MF$ smoothly crosses over from a maximal value of 1 to a minimal value of $2\gamma\p{n}$.
+The minimal value of $\bbk{\ss}_\MF$ approached as $h\to\infty$ can be explained with arguments identical to those in the paragraph containing Eq.~\eqref{eq:lim_ss}, which now imply that
+\begin{align}
+  \lim_{h\to\infty} \bbk{\ss}_\MF
   = \Tr\sp{\p{\diag\op{\xx}}^2}
@@
 \end{align}
-The vanishing magnetization $\sigma_\MF=0$ in Figure \ref{fig:mean_mag_int_XX} is protected by symmetries of $H_\spin$ and $\ket\XX$.
-For all initial states that we have considered, the value of $\bk{S_\z}=0$ is conserved by the spin Hamiltonian $H_\spin$.
-Moreover, both the spin Hamiltonian $H_\spin$ and the state $\ket\XX$ are invariant (up to global phase) under the action of $R_\z^\pi$, where $R_\z^\theta\equiv e^{-\i\theta S_\z}$.
+The vanishing magnetization $\sigma_\MF=0$ in Figure \ref{fig:mean_mag_int_XX} is protected by symmetries of $\H_\spin$ and $\ket\XX$.
+For all initial states that we have considered, the value of $\bk{\S_\z}=0$ is conserved by the spin Hamiltonian $\H_\spin$.
+Moreover, both the spin Hamiltonian $\H_\spin$ and the state $\ket\XX$ are invariant (up to global phase) under the action of $\R_\z^\pi$, where $\R_\z^\theta\equiv e^{-\i\theta \S_\z}$.
 Similarly to Eqs.~\eqref{eq:sym_x_start}--\eqref{eq:sym_x_end}, this symmetry implies that
 \begin{align}
-  &\bk{S_\x} = \bk{{R_\z^\pi}^\dag S_\x R_\z^\pi} = -\bk{S_\x} = 0, \\
-  &\bk{S_\y} = \bk{{R_\z^\pi}^\dag S_\y R_\z^\pi} = -\bk{S_\y} = 0
+  &\bk{\S_\x} = \bk{\R_\z^\pi{}^\dag \S_\x \R_\z^\pi} = -\bk{\S_\x} = 0, \\
+  &\bk{\S_\y} = \bk{\R_\z^\pi{}^\dag \S_\y \R_\z^\pi} = -\bk{\S_\y} = 0
 \end{align}
@@
 
-Turning now to mean-field results for the initial multi-cat state $\ket\XXI$ in Figure \ref{fig:mean_mag_int_XXI}, we remark that the magnetization $\sigma_\MF$ and interaction energy $\bbk{\ss}_\MF$ behave identically to those for the initial spin-polarized state $\ket\X$ in Figure \ref{fig:mean_mag_int_X}.
+Turning now to mean-field results for the initial kitten state $\ket\XXI$ in Figure \ref{fig:mean_mag_int_XXI}, we remark that the magnetization $\sigma_\MF$ and interaction energy $\bbk{\ss}_\MF$ behave identically to those for the initial spin-polarized state $\ket\X$ in Figure \ref{fig:mean_mag_int_X}.
 This finding can be understood through the fact that
 \begin{align}
-  \ket\XXI \simeq T_\z^{\pi/2} R_\z^{\pi/2} \ket\X,
-\end{align}
-where $T_\z^\theta\equiv e^{-\i\theta S_\z^2}$.
-The operators $R_\z^\theta$ and $T_\z^\theta$ are generated by axial fields that respect permutational symmetry, and therefore commute with the spin Hamiltonian $H_\spin$, so
-\begin{align}
-  e^{-\i t H_\spin} \ket\XXI
-  &\simeq e^{-\i t H_\spin} T_\z^{\pi/2} R_\z^{\pi/2} \ket\X \\
-  &\simeq T_\z^{\pi/2} R_\z^{\pi/2} e^{-\i t H_\spin} \ket\X.
+  \ket\XXI \simeq \T_\z^{\pi/2} \R_\z^{\pi/2} \ket\X,
+\end{align}
+where $\T_\z^\theta\equiv e^{-\i\theta \S_\z^2}$.
+The operators $\R_\z^\theta$ and $\T_\z^\theta$ are generated by axial fields that respect permutational symmetry, and therefore commute with the spin Hamiltonian $\H_\spin$, so
+\begin{align}
+  e^{-\i t \H_\spin} \ket\XXI
+  &\simeq e^{-\i t \H_\spin} \T_\z^{\pi/2} \R_\z^{\pi/2} \ket\X \\
+  &\simeq \T_\z^{\pi/2} \R_\z^{\pi/2} e^{-\i t \H_\spin} \ket\X.
 \end{align}
@@
 \begin{align}
-  {R_\z^\theta}^\dag {T_\z^\theta}^\dag \, \ss \, T_\z^\theta R_\z^\theta
+  \R_\z^\theta{}^\dag \T_\z^\theta{}^\dag \, \ss \, \T_\z^\theta \R_\z^\theta
   = \ss,
 \end{align}
-which implies that the interaction energy $\bk{\ss}$ throughout dynamics of the initial multi-cat state $\ket\XXI$ is the same as that of the spin-polarized state $\ket\X$.
+which implies that the interaction energy $\bk{\ss}$ throughout dynamics of the initial kitten state $\ket\XXI$ is the same as that of the spin-polarized state $\ket\X$.
 
@@
 That is, nonzero components $\bbk{\bar s_{\mu\nu}}_\MF$ of $\bbk{\bar{\v s}}_\MF$ always have $\mu=\pm\nu$ (see discussion below).
-\item \label{pt:twist} The twist operator $T_\z^\theta$ acts trivially on the diagonal and anti-diagonal components of $\bar{\v s}$, which together with point \ref{pt:diags} implies that $\bbk{{T_\z^\theta}^\dag \, \bar{\v s} \, T_\z^\theta}_\MF = \bbk{\bar{\v s}}_\MF$.
-\item \label{pt:rot} The rotation operator $R_\z^\theta$ merely rotates the magnetization vector $\bbk{\vec\sigma}_\MF$ without changing its magnitude.
+\item \label{pt:twist} The twist operator $\T_\z^\theta$ acts trivially on the diagonal and anti-diagonal components of $\bar{\v s}$, which together with point \ref{pt:diags} implies that $\bbk{\T_\z^\theta{}^\dag \, \bar{\v s} \, \T_\z^\theta}_\MF = \bbk{\bar{\v s}}_\MF$.
+\item \label{pt:rot} The rotation operator $\R_\z^\theta$ merely rotates the magnetization vector $\bbk{\vec\sigma}_\MF$ without changing its magnitude.
 \end{enumerate}
@@
   \sigma_\MF = \abs{\bbk{\vec\sigma}_\MF}
-  = \abs{\bbk{{R_\z^\theta}^\dag {T_\z^\theta}^\dag \,
-      \vec\sigma \, T_\z^\theta R_\z^\theta}_\MF}
+  = \abs{\bbk{\R_\z^\theta{}^\dag \T_\z^\theta{}^\dag \,
+      \vec\sigma \, \T_\z^\theta \R_\z^\theta}_\MF}
 \end{align}
@@
 This observation, nominally a numerical result of mean-field simulations, can be understood as follows.
-The eigenstates $\ket{m,w}$ of $H_\spin$ are uniquely identified by definite numbers $m=(m_s,m_{s-1},\cdots,m_{-s})$ of atoms occupying each internal spin state $\mu\in\set{s,s-1,\cdots,-s}$, and an auxiliary index $w$ that encodes how $\ket{m,w}$ transforms under permutations of all spins (see Appendix \ref{sec:pert_theory})\footnote{Seen otherwise, since $S_{\mu\mu}$ commutes with $H_\spin$, eigenvectors of $H_\spin$ can be indexed by eigenvalues of $S_{\mu\mu}$.
-The number $m_\mu$ is then the eigenvalue of $\ket{m,w}$ with respect to $S_{\mu\mu}$, i.e.~$S_{\mu\mu}\ket{m,w} = m_\mu\ket{m,w}$, while $w$ encodes all other information required to uniquely specify $\ket{m,w}$.}.
-The operator $\bar s_{\mu\nu}=\frac1N S_{\mu\nu}$ with $\mu\ne\nu$ couples the state $\ket{m,w}$ to states $\ket{m',w'}$ in which $(m_\mu',m'_\nu)=(m_\mu+1,m_\nu-1)$.
-Generically, states $\ket{m,w}$ and $\ket{m',w'}$ with $m\ne m'$ will have different energies, so their coherence averages to zero when evaluating time-averaged expectation values.
+The eigenstates $\ket{m,w}$ of $\H_\spin$ are uniquely identified by definite numbers $m=(m_s,m_{s-1},\cdots,m_{-s})$ of atoms occupying each internal spin state $\mu\in\set{s,s-1,\cdots,-s}$, and an auxiliary index $w$ that encodes how $\ket{m,w}$ transforms under permutations of all spins (see Appendix \ref{sec:pert_theory})\footnote{Seen otherwise, since $\S_{\mu\mu}$ commutes with $\H_\spin$, eigenvectors of $\H_\spin$ can be indexed by eigenvalues of $\S_{\mu\mu}$.
+The number $m_\mu$ is then the eigenvalue of $\ket{m,w}$ with respect to $\S_{\mu\mu}$, i.e.~$\S_{\mu\mu}\ket{m,w} = m_\mu\ket{m,w}$, while $w$ encodes all other information required to uniquely specify $\ket{m,w}$.}.
+The operator $\bar s_{\mu\nu}=\frac1N \S_{\mu\nu}$ with $\mu\ne\nu$ couples the state $\ket{m,w}$ to states $\ket{m',w'}$ in which $(m_\mu',m'_\nu)=(m_\mu+1,m_\nu-1)$.
+Generically, states $\ket{m,w}$ and $\ket{m',w'}$ with $m\ne m'$ will have different energies, so their coherence oscillates and averages to zero when evaluating time-averaged expectation values.
 
 However, degeneracies yield stationary (time-independent) coherences that survive time-averaging.
-In the weak-field limit $J\phi/U\to0$, such a degeneracy occurs at the mean-field level between PS states differing only in the populations $m_\mu,m_{-\mu}$ (with a fixed value of $m_\mu+m_{-\mu}$), as the effective Hamiltonian becomes $H_\MF^\eff\propto\sum_\mu\mu^2m_\mu$.
-This symmetry is preserved at all orders in perturbation theory\footnote{Only even powers of the ``perturbation'' $\sum_q\sin\p{q}s_{\z,q}$ can be nonzero within the PS manifold, and even powers of this perturbation exhibit the same mean-field degeneracy between states differing only in the populations $m_\mu,m_{-\mu}$.}, so some coherence between such states is preserved as $(J\phi/U)\to(J\phi/U)_\crit$, although this coherence decays as perturbative corrections to degenerate eigenstates cause them to leak out of the PS manifold (and thereby have a smaller overlap with the initial state $\ket\X$).
-Note that beyond-mean-field effects break the symmetry protecting anti-diagonal components of $\bbk{\v s}_\MF$, causing them to decay on time scales that should diverge as $N\to\infty$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions and outlooks}
+In the weak-field limit $h\to0$, such a degeneracy occurs at the mean-field level between PS states differing only in the populations $m_\mu,m_{-\mu}$ (with a fixed value of $m_\mu+m_{-\mu}$), as the effective Hamiltonian becomes $\H_\MF^\eff\propto\sum_\mu\mu^2m_\mu$.
+This symmetry is preserved at all orders in perturbation theory\footnote{Only even powers of the ``perturbation'' $\sum_q\sin\p{q}\s_{\z,q}$ can be nonzero within the PS manifold, and even powers of this perturbation exhibit the same mean-field degeneracy between states differing only in the populations $m_\mu,m_{-\mu}$.}, so some coherence between such states is preserved as $h\to h_\crit$, although this coherence decays as perturbative corrections to degenerate eigenstates cause them to leak out of the PS manifold (and thereby have a smaller overlap with the initial state $\ket\X$).
+Note that beyond-mean-field effects break the symmetry protecting anti-diagonal components of $\bbk{\v\s}_\MF$, causing them to decay on time scales that should diverge as $N\to\infty$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusions and future directions}
 \label{sec:conclusions}
@@
 We then considered the effect of spin-orbit coupling (SOC) induced by external driving lasers, finding that maintaining integrity of the spin model seems to require weak SOC, which in turn gives rise to a (synthetic) inhomogeneous magnetic field.
-Taking a closer look at the external controls that induce SOC, we found a simple three-laser drive that enables homogeneously addressing nuclear spins with a variety of spin Hamiltonians, enabling for example the preparation of ``multi-cat'' states in which each spin is in a coherent superposition of states spin-polarized in opposite directions.
+Taking a closer look at the external controls that induce SOC, we found a simple three-laser drive that enables homogeneously addressing nuclear spins with a variety of spin Hamiltonians, enabling for example the preparation of ``kitten'' states in which each spin is in a coherent superposition of states spin-polarized in opposite directions.
 Finally, we examined dynamical behavior of the SU($n$) spin model at the mean-field level, finding that long-time observables obey simple scaling relations with $n$, and that when $n>2$ dynamical behavior can be highly sensitive to intra-spin coherences.
@@
 Our work makes important progress in understanding the SU($n$) Fermi-Hubbard model in experimentally relevant parameter regimes, and we expect our findings to be readily testable in experiments with ultracold atoms.
-We hope this work provides a starting point for simulating SYK-like models in cold atomic platforms.
-In follow-up work, it would be interesting to study the the relationship between initial states and dynamical phases of our SU($n$) spin model more systematically, or to consider the effect of quantum corrections to mean-field behavior.
-There is also room to improve on the three-laser drive we described in this work, where it is natural to ask what ingredients and techniques are necessary to implement universal control of individual nuclear spins.
+Given the possibility for long-range SU($n$) interactions, we hope our work stimulates further efforts into simulating SY and SYK-like models \cite{sachdev1993gapless} in cold atomic platforms.
+In follow-up work, it would be interesting to study the the relationship between initial states and dynamical phases of our SU($n$) spin model more systematically, and to consider the effect of quantum corrections to mean-field behavior.
+There is also room to improve on the three-laser drive introduced in this work, for which it is natural to ask what additional techniques or ingredients are necessary to implement universal control of individual nuclear spins.
 Universal control would allow for an experimental study of $n$-dependence (including even/odd-$n$ parity effects) in a single experimental platform, simply by controlling the occupation and coherence of internal spin states.
 Finally, one can also study the SU($n$) Hubbard model in the super-exchange regime that gives rise to a real-space (as opposed to momentum-space) spin model, where SOC gives rise to chiral multilevel spin interactions.
-Unlike our present work, the super-exchange regime does not require weak SOC, and therefore has a larger parameter space in which to explore dynamical behaviors.
+Unlike our present work, the super-exchange regime does not require weak SOC, and therefore has a larger parameter space in which to explore dynamical behavior.
 
@@
 \begin{align}
-  H_0 = \sum_{i<j} g_{ij} \, \Pi_{ij},
-  &&
-  \Pi_{ij} \equiv \v s_i\c\v s_j
-  = \sum_{\mu,\nu} s_{\mu\nu i} s_{\nu\mu j},
+  \H_0 = \sum_{i<j} g_{ij} \, \hat\Pi_{ij},
+  &&
+  \hat\Pi_{ij} \equiv \v\s_i\cdot\v\s_j
+  = \sum_{\mu,\nu} \s_{\mu\nu i} \s_{\nu\mu j},
   \label{eq:H_0}
 \end{align}
-where $g_{ij}$ are (real) scalar coefficients for the permutation operators $\Pi_{ij}$, and $s_{\mu\nu i}\equiv c_{\mu i}^\dag c_{\nu i}$ is a transition operator for spin $i$.
+where $g_{ij}$ are (real) scalar coefficients for the permutation operators $\hat\Pi_{ij}$, and $\s_{\mu\nu i}\equiv \c_{\mu i}^\dag \c_{\nu i}$ is a transition operator for spin $i$.
 We can then consider the addition of, for example, an inhomogeneous magnetic field or Ising couplings,
 \begin{align}
-  H_{\t{field}} = \sum_i B_i s_{\z,i},
-  &&
-  H_{\t{Ising}} = \sum_{i\ne j} J_{ij} s_{\z,i} s_{\z,j},
+  \H_{\t{field}} = \sum_i B_i \s_{\z,i},
+  &&
+  \H_{\t{Ising}} = \sum_{i\ne j} J_{ij} \s_{\z,i} \s_{\z,j},
   \label{eq:field_ising}
@@
 or more generally an $M$-body operator\footnote{At face value, an $M$-body operator with $M>2$ does not typically appear in experiments.
-Nonetheless, considering $M>2$ illuminates the structure of eigenstates (and eigenvalues) of $H_0$, and allows us to go to high orders in perturbation theory with single- and two-body perturbations.}
-\begin{align}
-  \O\p{w,X} = \sum_{k\in\D_N\p{M}} w_k X_k,
+Nonetheless, considering $M>2$ illuminates the structure of eigenstates (and eigenvalues) of $\H_0$, and allows us to go to high orders in perturbation theory with single- and two-body perturbations.}
+\begin{align}
+  \O(w,\hat X) = \sum_{k\in\D_N\p{M}} w_k \hat X_k,
 \end{align}
@@
 coefficients $w_k\equiv w_{k_1k_2\cdots k_M}$; $X$ is an $M$-spin
-operator, e.g.~$s_\z\otimes s_\z$ in the case of Ising interactions
+operator, e.g.~$\s_\z\otimes \s_\z$ in the case of Ising interactions
 with $M=2$; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is a list of the
-individual spins $k_i\in\ZZ_N\equiv\set{1,2,\cdots,N}$ that the operator $X_k\equiv X_{k_1k_2\cdots k_M}$ acts on; and
+individual spins $k_i\in\ZZ_N\equiv\set{1,2,\cdots,N}$ that the operator $\hat X_k\equiv \hat X_{k_1k_2\cdots k_M}$ acts on; and
 \begin{align}
@@
 \end{align}
-is the strictly ``off-diagonal'' part of $\ZZ_N^M$, which is necessary to identify for a consistent definition of $X_k$ as an $M$-body operator.
-In this notation, the magnetic field and Ising Hamiltonians in Eq.~\eqref{eq:field_ising} respectively become $\O\p{B,s_{\z}}$ and $\O\p{J,s_\z\otimes s_\z}$.
-
-If the addition $\O\p{w,X}$ to the SU($n$)-symmetric Hamiltonian $H_0$ in Eq.~\eqref{eq:H_0} is sufficiently small, namely with operator norm $\norm{\O\p{w,X}}$ less than half the spectral gap $\Delta_{\t{gap}}$ of $H_0$, $\norm{\O\p{w,X}}<\Delta_{\t{gap}}/2$, then we can treat the effect of $\O\p{w,X}$ on the ground-state PS manifold $\E_0$ perturbatively.
-The effective Hamiltonians $H_\eff^{(1)}$ and $H_\eff^{(2)}$ induced by $\O\p{w,X}$ on the PS manifold $\E_0$ at leading orders in perturbation theory are \cite{bravyi2011schrieffer}
-\begin{align}
-  H_\eff^{(1)} = \P_0 \O\p{w,X} \P_0,
-  &&
-  H_\eff^{(2)} = - \sum_{\Delta\ne0}
-  \f1\Delta \P_0 \O\p{w,X} \P_\Delta \O\p{w,X} \P_0,
+is the strictly ``off-diagonal'' part of $\ZZ_N^M$, which is necessary to identify for a consistent definition of $\hat X_k$ as an $M$-body operator.
+In this notation, the magnetic field and Ising Hamiltonians in Eq.~\eqref{eq:field_ising} respectively become $\O\p{B,\s_{\z}}$ and $\O\p{J,\s_\z\otimes \s_\z}$.
+
+If the addition $\O(w,\hat X)$ to the SU($n$)-symmetric Hamiltonian $\H_0$ in Eq.~\eqref{eq:H_0} is sufficiently small, namely with operator norm $\norm*{\O(w,\hat X)}$ less than half the spectral gap $\Delta_{\t{gap}}$ of $\H_0$, $\norm*{\O(w,\hat X)}<\Delta_{\t{gap}}/2$, then we can treat the effect of $\O(w,\hat X)$ on the ground-state PS manifold $\E_0$ perturbatively.
+The effective Hamiltonians $\H_\eff^{(1)}$ and $\H_\eff^{(2)}$ induced by $\O(w,\hat X)$ on the PS manifold $\E_0$ at leading orders in perturbation theory are \cite{bravyi2011schrieffer}
+\begin{align}
+  \H_\eff^{(1)} = \P_0 \O(w,\hat X) \P_0,
+  &&
+  \H_\eff^{(2)} = - \sum_{\Delta\ne0}
+  \f1\Delta \P_0 \O(w,\hat X) \P_\Delta \O(w,\hat X) \P_0,
   \label{eq:H_eff}
 \end{align}
-where $\P_\Delta$ is a projector onto the eigenspace $\E_\Delta$ of $H_0$ with interaction energy $\Delta$ above that of the PS manifold.
-The first order effective Hamiltonian $H_\eff^{(1)}$ simply projects $\O\p{w,X}$ onto the PS manifold $\E_0$, and takes the form
-\begin{align}
-  H_\eff^{(1)} = \mean w \, \col X,
+where $\P_\Delta$ is a projector onto the eigenspace $\E_\Delta$ of $\H_0$ with interaction energy $\Delta$ above that of the PS manifold.
+The first order effective Hamiltonian $\H_\eff^{(1)}$ simply projects $\O(w,\hat X)$ onto the PS manifold $\E_0$, and takes the form
+\begin{align}
+  \H_\eff^{(1)} = \mean w \, \col X,
   \label{eq:H_eff_1}
@@
   &&
-  \col X \equiv \sum_{k\in\D_N\p{M}} X_k,
+  \col X \equiv \sum_{k\in\D_N\p{M}} \hat X_k,
   \label{eq:mean_col}
@@
 with $\abs{\D_N\p{M}}=\prod_{j=0}^{M-1}\p{N-j}$.
-In the case of a magnetic field $s_\z$ or Ising interactions  $s_\z\otimes s_\z$, for example,
-\begin{align}
-  \col{s_\z} = \sum_i s_\z^{(i)} = S_\z,
-  &&
-  \col{s_\z\otimes s_\z}
-  = \sum_{i\ne j} s_\z^{(i)} s_\z^{(j)}
-  = S_\z^2 - N \sum_i s_{\z,i}^2.
-\end{align}
-The second order effective Hamiltonian $H_\eff^{(2)}$ in Eq.~\eqref{eq:H_eff} takes more work to simplify due to the presence of a projector $\P_\Delta$ onto the manifold $\E_\Delta$ of states with excitation energy $\Delta$.
-This projector essentially picks off the part of $\O\p{w,X}$ that is strictly off-diagonal with respect to the ground- and excited-state manifolds $\E_0$ and $\E_\Delta$.
-We therefore need to decompose $\O\p{w,X}$ into components that generate states of definite excitation energy when acting on PS states $\ket\psi\in\E_0$.
-The SU($n$) symmetry of $H_0$ enables such a decomposition to take the form
-\begin{align}
-  H_0 \O\p{w,X} \ket\psi
-  = \sum_\Delta \p{E_0+\Delta} \O\p{w^\Delta,X} \ket\psi,
+In the case of a magnetic field $\s_\z$ or Ising interactions  $\s_\z\otimes \s_\z$, for example,
+\begin{align}
+  \col{\s_\z} = \sum_i \s_\z^{(i)} = \S_\z,
+  &&
+  \col{\s_\z\otimes \s_\z}
+  = \sum_{i\ne j} \s_\z^{(i)} \s_\z^{(j)}
+  = \S_\z^2 - N \sum_i \s_{\z,i}^2.
+\end{align}
+The second order effective Hamiltonian $\H_\eff^{(2)}$ in Eq.~\eqref{eq:H_eff} takes more work to simplify due to the presence of a projector $\P_\Delta$ onto the manifold $\E_\Delta$ of states with excitation energy $\Delta$.
+This projector essentially picks off the part of $\O(w,\hat X)$ that is strictly off-diagonal with respect to the ground- and excited-state manifolds $\E_0$ and $\E_\Delta$.
+We therefore need to decompose $\O(w,\hat X)$ into components that generate states of definite excitation energy when acting on PS states $\ket\psi\in\E_0$.
+The SU($n$) symmetry of $\H_0$ enables such a decomposition to take the form
+\begin{align}
+  \H_0 \O(w,\hat X) \ket\psi
+  = \sum_\Delta \p{E_0+\Delta} \O(w^\Delta,\hat X) \ket\psi,
   &&
@@
 
-Equipped with the decomposition $\O\p{w,X}=\sum_\Delta\O\p{w^\Delta,X}$ with terms $\O\p{w^\Delta,X}$ that generate states of definite excitation energy $\Delta$, we can expand
-\begin{align}
-  H_{\t{eff}}^{(2)} = -\sum_{\Delta\ne0} \f1\Delta
-  \P_0 \O\p{w^\Delta,X}^2 \P_0.
+Equipped with the decomposition $\O(w,\hat X)=\sum_\Delta\O(w^\Delta,\hat X)$ with terms $\O(w^\Delta,\hat X)$ that generate states of definite excitation energy $\Delta$, we can expand
+\begin{align}
+  \H_{\t{eff}}^{(2)} = -\sum_{\Delta\ne0} \f1\Delta
+  \P_0 \O(w^\Delta,\hat X)^2 \P_0.
 \end{align}
@@
 \begin{align}
-  H_{\t{eff}}^{(2)}
-  = \sum_{\Delta\ne0} \f{w^\Delta\c w^\Delta}{N\p{N-1}\Delta}
+  \H_{\t{eff}}^{(2)}
+  = \sum_{\Delta\ne0} \f{w^\Delta\cdot w^\Delta}{N\p{N-1}\Delta}
   \p{\col{X}^2 - N \col{X^2}},
 \end{align}
-and if furthermore all $g_{ij}=-U/N$, as for $H_{\t{int}}$ in Eq.~\eqref{eq:H_int}, then the only relevant excitation energy is $\Delta=U$ (see Section \ref{sec:spin_wave}), and
-\begin{align}
-  w^U\c w^U = \sum_i \p{w_i-\mean w}^2 = N \widetilde{w}^2
+and if furthermore all $g_{ij}=-U/N$, as for $\H_{\t{int}}$ in Eq.~\eqref{eq:H_int}, then the only relevant excitation energy is $\Delta=U$ (see Section \ref{sec:spin_wave}), and
+\begin{align}
+  w^U\cdot w^U = \sum_i \p{w_i-\mean w}^2 = N \widetilde{w}^2
 \end{align}
@@
 \begin{align}
-  H_{\t{eff}}^{(2)}
+  \H_{\t{eff}}^{(2)}
   = \f{\widetilde{w}^2}{\p{N-1}U} \p{\col{X}^2 - N \col{X^2}}.
@@
 
-Here we construct the matrix $g^{(M)}$ that enables decomposing $M$-body operators $\O\p{w,X}$ into terms $\O\p{w^\Delta,X}$ that generate states of definite excitation energy $\Delta$ above the PS manifold, as in Eq.~\eqref{eq:excitation}.
+Here we construct the matrix $g^{(M)}$ that enables decomposing $M$-body operators $\O(w,\hat X)$ into terms $\O(w^\Delta,\hat X)$ that generate states of definite excitation energy $\Delta$ above the PS manifold, as in Eq.~\eqref{eq:excitation}.
 We work through the calculation of $g^{(1)}$ explicitly, and provide the result for $g^{(M)}$ from a generalized version of the same calculation.
-To this end, we consider the action of a single-body operator $\O\p{w,X}=\sum_iw_iX_i$ on an arbitrary PS state $\ket\psi\in\E_0$ and expand
-\begin{align}
-  H_0 \O\p{w,X} \ket\psi
-  = \f12 \sum_{i\ne j} \sum_k g_{ij} w_k \Pi_{ij} X_k \ket\psi,
+To this end, we consider the action of a single-body operator $\O(w,\hat X)=\sum_iw_i\hat X_i$ on an arbitrary PS state $\ket\psi\in\E_0$ and expand
+\begin{align}
+  \H_0 \O(w,\hat X) \ket\psi
+  = \f12 \sum_{i\ne j} \sum_k g_{ij} w_k \hat\Pi_{ij} \hat X_k \ket\psi,
   \label{eq:single_body_diagnosis_start}
@@
 The sum in Eq.~\eqref{eq:single_body_diagnosis_start} has terms with $k\in\set{i,j}$ and terms with $k\notin\set{i,j}$.
-In the case of $k\notin\set{i,j}$, the permutation operator $\Pi_{ij}$ commutes with $X_k$ and annihilates on $\ket\psi$, and we can replace the sum
+In the case of $k\notin\set{i,j}$, the permutation operator $\hat\Pi_{ij}$ commutes with $\hat X_k$ and annihilates on $\ket\psi$, and we can replace the sum
 \begin{align}
@@
   \f12 \sum_{i\ne j} \sum_{k\notin\set{i,j}}
-  g_{ij} w_k \Pi_{ij} X_k \ket\psi
-  = E_0 \O\p{w,X} \ket\psi
-  - \f12 \sum_{i\ne j} \sum_{k\in\set{i,j}} g_{ij} w_k X_k \ket\psi,
+  g_{ij} w_k \hat\Pi_{ij} \hat X_k \ket\psi
+  = E_0 \O(w,\hat X) \ket\psi
+  - \f12 \sum_{i\ne j} \sum_{k\in\set{i,j}} g_{ij} w_k \hat X_k \ket\psi,
 \end{align}
@@
   \f12 \sum_{i\ne j} \sum_{k\notin\set{i,j}}
-  g_{ij} w_k \Pi_{ij} X_k \ket\psi
-  = E_0 \O\p{w,X} \ket\psi - \sum_k g_k w_k X_k \ket\psi.
+  g_{ij} w_k \hat\Pi_{ij} \hat X_k \ket\psi
+  = E_0 \O(w,\hat X) \ket\psi - \sum_k g_k w_k \hat X_k \ket\psi.
 \end{align}
@@
   \f12 \sum_{\substack{i\ne j\\k\in\set{i,j}}}
-  g_{ij} w_k \Pi_{ij} X_k \ket\psi
-  = \sum_{i,j} g_{ij} w_j X_i \ket\psi.
+  g_{ij} w_k \hat\Pi_{ij} \hat X_k \ket\psi
+  = \sum_{i,j} g_{ij} w_j \hat X_i \ket\psi.
 \end{align}
@@
 \begin{align}
-  H_0 \O\p{w,X} \ket\psi
-  = E_0 \O\p{w,X} \ket\psi
-  + \sum_k \sp{\sum_j g_{kj} w_j - g_k w_k} X_k \ket\psi.
-\end{align}
-The action of the single-body perturbation $\O\p{w,X}$ on a permutationally symmetric state therefore generates an eigenstate of $H_0$ with interaction energy $E_0+\Delta$ if the vector $w=\sum_k w_k \ket{k}$ satisfies the eigenvalue equation
-\begin{align}
-  g^{(1)} \c w = \Delta w, && g^{(1)} \equiv g - \diag\vec g,
+  \H_0 \O(w,\hat X) \ket\psi
+  = E_0 \O(w,\hat X) \ket\psi
+  + \sum_k \sp{\sum_j g_{kj} w_j - g_k w_k} \hat X_k \ket\psi.
+\end{align}
+The action of the single-body perturbation $\O(w,\hat X)$ on a permutationally symmetric state therefore generates an eigenstate of $\H_0$ with interaction energy $E_0+\Delta$ if the vector $w=\sum_k w_k \ket{k}$ satisfies the eigenvalue equation
+\begin{align}
+  g^{(1)} \cdot w = \Delta w, && g^{(1)} \equiv g - \diag\vec g,
   \label{eq:single_body_eig}
@@
 \begin{align}
-  g^{(M)} \c w = \Delta w,
+  g^{(M)} \cdot w = \Delta w,
   \label{eq:multi_body_problem}
@@
 
-If the interaction Hamiltonian $H_0$ is translationally invariant, then the single-body eigenvalue problem in Eq.~\eqref{eq:single_body_eig} is solvable analytically.
+If the interaction Hamiltonian $\H_0$ is translationally invariant, then the single-body eigenvalue problem in Eq.~\eqref{eq:single_body_eig} is solvable analytically.
 In this case, the couplings $g_{ij}$ depend only on the separation $\abs{i-j}$, so eigenvectors of $g$ are plane waves of the form
 \begin{align}
-  w_k \equiv \sum_{d\in\ZZ_L^D} e^{\i d\c k} \ket{d},
+  w_k \equiv \sum_{d\in\ZZ_L^D} e^{\i d\cdot k} \ket{d},
 \end{align}
@@
 \begin{align}
-  g \c w_k
-  = \sum_{c,d\in\ZZ_L^D} g_{cd} e^{\i d\c k} \ket{c}
-  = \sum_{c,d\in\ZZ_L^D} g_{c,c+d} e^{\i\p{c+d}\c k} \ket{c}
-  = \sum_{d\in\ZZ_L^D} g_{0,d} \cos\p{d\c k} w_k,
+  g \cdot w_k
+  = \sum_{c,d\in\ZZ_L^D} g_{cd} e^{\i d\cdot k} \ket{c}
+  = \sum_{c,d\in\ZZ_L^D} g_{c,c+d} e^{\i\p{c+d}\cdot k} \ket{c}
+  = \sum_{d\in\ZZ_L^D} g_{0,d} \cos\p{d\cdot k} w_k,
 \end{align}
@@
 \begin{align}
-  g^{(1)} \c w_k = \Delta_k w_k,
-  &&
-  \Delta_k \equiv \sum_{d\in\ZZ_L^D} g_{0,d} \sp{\cos\p{d\c k}-1},
+  g^{(1)} \cdot w_k = \Delta_k w_k,
+  &&
+  \Delta_k \equiv \sum_{d\in\ZZ_L^D} g_{0,d} \sp{\cos\p{d\cdot k}-1},
 \end{align}
@@
   \ket{a} = \f1{\sqrt{\C\p{a}}}
-  \sum_{\substack{\t{distinct}\\\t{permutations}\\\Pi\,\t{of}\,\tilde a}}
-  \Pi\ket{\tilde a},
+  \sum_{\substack{\t{distinct}\\\t{permutations}\\\hat\Pi\,\t{of}\,\tilde a}}
+  \hat\Pi\ket{\tilde a},
   &&
@@
 \begin{multline}
-  \Delta \times H_{\t{3LD}}^{\t{single}}
-  = \p{\Omega_+^2 - \Omega_-^2} s_\z
+  \Delta \times \H_{\t{3LD}}^{\t{single}}
+  = \p{\Omega_+^2 - \Omega_-^2} \s_\z
   - \sum_{\sigma\in\set{\pm1}} \Omega_0 \Omega_\sigma
-  \p{\sigma s_{\tilde\eta_\sigma,\x}
-    + s_\z s_{\tilde\eta_\sigma,\x}
-    + s_{\tilde\eta_\sigma,\x} s_\z} \\
-  - \Omega_0^2 s_\z^2 - \p{\Omega_+ + \Omega_-}^2 s_{\tilde\eta_0,\x}^2
-  - \p{\Omega_+ - \Omega_-}^2 s_{\tilde\eta_0,\y}^2,
+  \p{\sigma \s_{\tilde\eta_\sigma,\x}
+    + \s_\z \s_{\tilde\eta_\sigma,\x}
+    + \s_{\tilde\eta_\sigma,\x} \s_\z} \\
+  - \Omega_0^2 \s_\z^2 - \p{\Omega_+ + \Omega_-}^2 \s_{\tilde\eta_0,\x}^2
+  - \p{\Omega_+ - \Omega_-}^2 \s_{\tilde\eta_0,\y}^2,
 \end{multline}
-where $s_{\eta\alpha} \equiv e^{-\i\eta s_\z} s_\alpha e^{\i\eta s_\z}$ is a rotated spin-$\alpha$ operator (with e.g.~$s_{\pi/2,\x}=s_\y$), and
+where $\s_{\eta\alpha} \equiv e^{-\i\eta \s_\z} \s_\alpha e^{\i\eta \s_\z}$ is a rotated spin-$\alpha$ operator (with e.g.~$\s_{\pi/2,\x}=\s_\y$), and
 \begin{align}
@@
 \begin{align}
-  H_\spin = -\f{U}{2N}\v S\c\v S + J\phi \sum_q \sin\p{q} s_{\z,q}
+  \H_\spin = -\f{u}{2N}\v\S\cdot\v\S + 2J\phi \sum_q \sin\p{q} \s_{\z,q}
 \end{align}
 in Eq.~\eqref{eq:H_spin} of the main text.
-We begin by decomposing individual spin operators into Schwinger bosons as $s_{\mu\nu q} = b_{\mu q}^\dag b_{\nu q}$, such that the spin Hamiltonian becomes
-\begin{align}
-  H_\spin \to H_{\t{boson}}
-  = -\f{U}{2N} \sum_{p,q,\mu,\nu}
-  b_{\mu p}^\dag b_{\nu p} b_{\nu q}^\dag b_{\mu q}
-  + J\phi \sum_{q,\mu} \sin\p{q} \mu\, b_{\mu q}^\dag b_{\mu q}.
+We begin by decomposing individual spin operators into Schwinger bosons as $\s_{\mu\nu q} = \b_{\mu q}^\dag \b_{\nu q}$, such that the spin Hamiltonian becomes
+\begin{align}
+  \H_\spin \to \H_{\t{boson}}
+  = -\f{u}{2N} \sum_{p,q,\mu,\nu}
+  \b_{\mu p}^\dag \b_{\nu p} \b_{\nu q}^\dag \b_{\mu q}
+  + 2J\phi \sum_{q,\mu} \sin\p{q} \mu\, \b_{\mu q}^\dag \b_{\mu q}.
 \end{align}
@@
 \begin{align}
-  \i \partial_t b_{\mu q}
-  = -\f{U}{N} \sum_{\nu,p} b_{\nu p}^\dag b_{\mu p} b_{\nu q}
-  + J\phi \sin\p{q} \mu\, b_{\mu q}.
-\end{align}
-Our mean-field theory then treats all boson operators in these equations of motion as complex numbers, $b_{\mu q}\to\bk{b_{\mu q}}_\MF$, with the initial value $\bk{b_{\mu q}\p{t=0}}_\MF$ equal to the initial amplitude of spin $q$ in state $\mu$.
-Specifically, for an $N$-fold product state of the form $\ket\psi=\bigotimes_q\sum_\mu\psi_{\mu q}\ket{\mu}$ we set $\bk{b_{\mu q}\p{t=0}}_\MF = \psi_{\mu q}$.
-For pure initial product states, this mean-field treatment of the boson operators $b_{\mu q}$ is mathematically equivalent to a mean-field treatment of the spin operators $s_{\mu\nu q}$, as in Eq.~\eqref{eq:H_MF}, but reduces the number of variables to keep track of by a factor of $\sim n$.
+  \i \partial_t \b_{\mu q}
+  = -\f{u}{N} \sum_{\nu,p} \b_{\nu p}^\dag \b_{\mu p} \b_{\nu q}
+  + 2J\phi \sin\p{q} \mu\, \b_{\mu q}.
+\end{align}
+Our mean-field theory then treats all boson operators in these equations of motion as complex numbers, $\b_{\mu q}\to\bk{\b_{\mu q}}_\MF$, with the initial value $\bk{\b_{\mu q}\p{t=0}}_\MF$ equal to the initial amplitude of spin $q$ in state $\mu$.
+Specifically, for an $N$-fold product state of the form $\ket\psi=\bigotimes_q\sum_\mu\psi_{\mu q}\ket{\mu}$ we set $\bk{\b_{\mu q}\p{t=0}}_\MF = \psi_{\mu q}$.
+For pure initial product states, this mean-field treatment of the boson operators $\b_{\mu q}$ is mathematically equivalent to a mean-field treatment of the spin operators $\s_{\mu\nu q}$, as in Eq.~\eqref{eq:H_MF}, but reduces the number of variables to keep track of by a factor of $\sim n$.
 
@@
 \begin{align}
-  H = \sum_{\substack{\mu,\nu,\rho,\sigma\\j<k}}
-  g^{\mu\nu j}_{\rho\sigma k} s_{\mu\nu j} s_{\rho\sigma k}
-  + \sum_{\mu,\nu,j} \epsilon_{\mu\nu j} s_{\mu\nu j},
+  \H = \sum_{\substack{\mu,\nu,\rho,\sigma\\j<k}}
+  g^{\mu\nu j}_{\rho\sigma k} \s_{\mu\nu j} \s_{\rho\sigma k}
+  + \sum_{\mu,\nu,j} \epsilon_{\mu\nu j} \s_{\mu\nu j},
   \label{eq:quadratic_spin}
 \end{align}
-where $\mu,\nu$ index orthogonal states of an $n$-level spin; $j,k$ index one of $N$ spins; $g^{\mu\nu j}_{\rho\sigma k}$ and $\epsilon_{\mu\nu j}$ are scalars; and $s_{\mu\nu j}=\op{\mu}{\nu}_j$ is a transition operator for spin $j$.
+where $\mu,\nu$ index orthogonal states of an $n$-level spin; $j,k$ index one of $N$ spins; $g^{\mu\nu j}_{\rho\sigma k}$ and $\epsilon_{\mu\nu j}$ are scalars; and $\s_{\mu\nu j}=\op{\mu}{\nu}_j$ is a transition operator for spin $j$.
 Strictly speaking, Eq.~\eqref{eq:quadratic_spin} only defines the couplings $g^{\mu\nu j}_{\rho\sigma k}$ for $j<k$, so we enforce $g^{\mu\nu k}_{\rho\sigma j}=g^{\mu\nu j}_{\rho\sigma k}$ and $g^{\mu\nu j}_{\rho\sigma j}=0$ for completion.
-Decomposing spin operators into Schwinger bosons as $s_{\mu\nu j}=b_{\mu j}^\dag b_{\nu j}$, where $b_{\nu j}$ a annihilates a boson of type $\nu$ on site $j$, we can write this Hamiltonian as
-\begin{align}
-  H = \sum_{\substack{\mu,\nu,\rho,\sigma\\j<k}}
+Decomposing spin operators into Schwinger bosons as $\s_{\mu\nu j}=\b_{\mu j}^\dag \b_{\nu j}$, where $\b_{\nu j}$ a annihilates a boson of type $\nu$ on site $j$, we can write this Hamiltonian as
+\begin{align}
+  \H = \sum_{\substack{\mu,\nu,\rho,\sigma\\j<k}}
   g^{\mu\nu j}_{\rho\sigma k}
-  b_{\mu j}^\dag b_{\nu j} b_{\rho k}^\dag b_{\sigma k}
-  + \sum_{\mu,\nu,j} \epsilon_{\mu\nu j} b_{\mu j}^\dag b_{\nu j}.
+  \b_{\mu j}^\dag \b_{\nu j} \b_{\rho k}^\dag \b_{\sigma k}
+  + \sum_{\mu,\nu,j} \epsilon_{\mu\nu j} \b_{\mu j}^\dag \b_{\nu j}.
 \end{align}
@@
 \begin{align}
-  \i \partial_t b_{\alpha\ell} = \sp{b_{\alpha\ell}, H}
+  \i \partial_t \b_{\alpha\ell} = \sp{\b_{\alpha\ell}, \H}
   &= \sum_{\substack{\mu,\nu,\rho,\sigma\\j<k}}
   g^{\mu\nu j}_{\rho\sigma k}
-  \sp{b_{\alpha\ell}, b_{\mu j}^\dag b_{\nu j} b_{\rho k}^\dag b_{\sigma k}}
+  \sp{\b_{\alpha\ell}, \b_{\mu j}^\dag \b_{\nu j} \b_{\rho k}^\dag \b_{\sigma k}}
   + \sum_{\mu,\nu,j} \epsilon_{\mu\nu j}
-  \sp{b_{\alpha\ell}, b_{\mu j}^\dag b_{\nu j}} \\
+  \sp{\b_{\alpha\ell}, \b_{\mu j}^\dag \b_{\nu j}} \\
   &= \sum_{\mu,\nu,\rho,\sigma,k} g^{\mu\nu\ell}_{\rho\sigma k}
-  \sp{b_{\alpha\ell}, b_{\mu\ell}^\dag b_{\nu\ell}}
-  b_{\rho k}^\dag b_{\sigma k}
+  \sp{\b_{\alpha\ell}, \b_{\mu\ell}^\dag \b_{\nu\ell}}
+  \b_{\rho k}^\dag \b_{\sigma k}
   + \sum_{\mu,\nu} \epsilon_{\mu\nu\ell}
-  \sp{b_{\alpha\ell}, b_{\mu\ell}^\dag b_{\nu\ell}} \\
+  \sp{\b_{\alpha\ell}, \b_{\mu\ell}^\dag \b_{\nu\ell}} \\
   &= \sum_{\mu,\nu} \p{\sum_{\rho,\sigma,k}
-    g^{\mu\nu\ell}_{\rho\sigma k} b_{\rho k}^\dag b_{\sigma k}
+    g^{\mu\nu\ell}_{\rho\sigma k} \b_{\rho k}^\dag \b_{\sigma k}
     + \epsilon_{\mu\nu\ell}}
-  \sp{b_{\alpha\ell}, b_{\mu\ell}^\dag b_{\nu\ell}}
+  \sp{\b_{\alpha\ell}, \b_{\mu\ell}^\dag \b_{\nu\ell}}
 \end{align}
@@
 \begin{align}
-  \sp{b_{\alpha\ell}, b_{\mu\ell}^\dag b_{\nu\ell}}
-  = \delta_{\alpha\mu} \delta_{\alpha\nu} b_{\alpha\ell}
-  + \delta_{\alpha\mu} \p{1-\delta_{\alpha\nu}} b_{\nu\ell}
-  = \delta_{\alpha\mu} b_{\nu\ell},
+  \sp{\b_{\alpha\ell}, \b_{\mu\ell}^\dag \b_{\nu\ell}}
+  = \delta_{\alpha\mu} \delta_{\alpha\nu} \b_{\alpha\ell}
+  + \delta_{\alpha\mu} \p{1-\delta_{\alpha\nu}} \b_{\nu\ell}
+  = \delta_{\alpha\mu} \b_{\nu\ell},
 \end{align}
@@
 \begin{align}
-  \i \partial_t b_{\alpha\ell}
+  \i \partial_t \b_{\alpha\ell}
   = \sum_\nu \p{\sum_{\rho,\sigma,k}
-    g^{\alpha\nu\ell}_{\rho\sigma k} b_{\rho k}^\dag b_{\sigma k}
-    + \epsilon_{\alpha\nu\ell}} b_{\nu\ell}.
-\end{align}
-In the case of uniform SU($n$)-symmetric interactions of the form $\frac{g}{2}\v S\c\v S$ and a diagonal external field, we have
+    g^{\alpha\nu\ell}_{\rho\sigma k} \b_{\rho k}^\dag \b_{\sigma k}
+    + \epsilon_{\alpha\nu\ell}} \b_{\nu\ell}.
+\end{align}
+In the case of uniform SU($n$)-symmetric interactions of the form $\frac{g}{2}\v\S\cdot\v\S$ and a diagonal external field, we have
 \begin{align}
@@
 \begin{align}
-  \i \partial_t b_{\alpha\ell}
-  = g \sum_{\nu,k} b_{\nu k}^\dag b_{\alpha k} b_{\nu\ell}
-  + \epsilon_{\alpha\ell} b_{\alpha\ell}.
+  \i \partial_t \b_{\alpha\ell}
+  = g \sum_{\nu,k} \b_{\nu k}^\dag \b_{\alpha k} \b_{\nu\ell}
+  + \epsilon_{\alpha\ell} \b_{\alpha\ell}.
 \end{align}
@@
 \begin{align}
-  H_\spin
-  = -\f{U}{2N} \sum_{\mu,\nu} S_{\mu\nu} S_{\nu\mu}
-  + J\phi \sum_q \sin\p{q} s_{\z,q},
-\end{align}
-where $S_{\mu\nu} = \sum_q s_{\mu\nu q}$.
+  \H_\spin
+  = -\f{u}{2N} \sum_{\mu,\nu} \S_{\mu\nu} \S_{\nu\mu}
+  + 2J\phi \sum_q \sin\p{q} \s_{\z,q},
+\end{align}
+where $\S_{\mu\nu} = \sum_q \s_{\mu\nu q}$.
 The single-body operators that appear in this Hamiltonian have squared norms
 \begin{align}
-  \tr\p{s_{\mu\nu q}^\dag s_{\mu\nu q}} = 1
+  \tr\p{\s_{\mu\nu q}^\dag \s_{\mu\nu q}} = 1
   &&
@@
   &&
-  \tr\p{s_{\z,q}^\dag s_{\z,q}}
+  \tr\p{\s_{\z,q}^\dag \s_{\z,q}}
   = \sum_\mu \mu^2
@@
 \end{align}
-The Lax formulation (following Refs.~\cite{yuzbashyan2005nonequilibrium, yuzbashyan2006dynamical, yuzbashyan2006relaxation, yuzbashyan2015quantum, smale2019observation}) requires all single-body operators involved to have the same normalization, so we substitute $s_{\tilde\z,q}\equiv s_{\z,q}/\eta$ to expand
-\begin{align}
-  \f{H_\spin}{U}
-  = -\f1{2N} \sum_{\mu,\nu} S_{\mu\nu} S_{\nu\mu}
-  + \eta h \sum_q \sin\p{q} s_{\tilde\z,q},
+The Lax formulation (following Refs.~\cite{yuzbashyan2005nonequilibrium, yuzbashyan2006dynamical, yuzbashyan2006relaxation, yuzbashyan2015quantum, smale2019observation}) requires all single-body operators involved to have the same normalization, so we substitute $\s_{\tilde\z,q}\equiv \s_{\z,q}/\eta$ to expand
+\begin{align}
+  \f{\H_\spin}{u}
+  = -\f1{2N} \sum_{\mu,\nu} \S_{\mu\nu} \S_{\nu\mu}
+  + \eta h \sum_q \sin\p{q} \s_{\tilde\z,q},
   &&
@@
   &&
-  h \equiv \f{J\phi}{U}.
-\end{align}
-The normalized, dimensionless, $(n^2-1)$-component Lax vector $\vec\ell\p{z}$ associated with $H_\spin$, which is defined with an auxiliary complex parameter $z$, has components
+  h \equiv \f{2J\phi}{u}.
+\end{align}
+The normalized, dimensionless, $(n^2-1)$-component Lax vector $\vec\ell\p{z}$ associated with $\H_\spin$, which is defined with an auxiliary complex parameter $z$, has components
 \begin{align}
   \ell_\alpha\p{z}
-  = \f1N \sum_q \f{s_{\alpha,q}}{z-\sin q}
+  = \f1N \sum_q \f{\s_{\alpha,q}}{z-\sin q}
   + \delta_{\alpha,\tilde\z} \, \eta h,
 \end{align}
-where $\alpha$ indexes elements of a basis $\set{s_\alpha}$ of self-adjoint generators of SU($n$), with normalization $\tr\p{s_\alpha^2}=1$.
-The squared magnitude $\vec\ell\p{z}^2=\sum_\alpha\ell_\alpha\p{z}^2$ is a constant of motion (for any $z$), and its residues provide $N$ mutually commuting quantities whose weighted sum recovers $H_\spin$.
+where $\alpha$ indexes elements of a basis $\set{\s_\alpha}$ of self-adjoint generators of SU($n$), with normalization $\tr\p{\s_\alpha^2}=1$.
+The squared magnitude $\vec\ell\p{z}^2=\sum_\alpha\ell_\alpha\p{z}^2$ is a constant of motion (for any $z$), and its residues provide $N$ mutually commuting quantities whose weighted sum recovers $\H_\spin$.
 When $n=2$, conservation of these residues provides sufficient dynamical constraints to make the spin system fully integrable.
-In this case, dynamical behavior is governed by the roots of $\vec\ell\p{z}^2$, and the presence (or absence) of complex roots marks distinct dynamical phases of $H_\spin$.
+In this case, dynamical behavior is governed by the roots of $\vec\ell\p{z}^2$, and the presence (or absence) of complex roots marks distinct dynamical phases of $\H_\spin$.
 However, the size of Hilbert space grows with $n$, while the number of conserved quantities provided by the Lax analysis (namely, $N$) does not.
@@
 
-Within the permutationally symmetric manifold, we can replace $s_{\alpha,q}\to\bar s_\alpha\equiv\frac1N\sum_q s_{\alpha,q}$ at the cost of $O(1/N)$ errors that vanish as $N\to\infty$, so taking this limit we find
+Within the permutationally symmetric manifold, we can replace $\s_{\alpha,q}\to\bar s_\alpha\equiv\frac1N\sum_q \s_{\alpha,q}$ at the cost of $O(1/N)$ errors that vanish as $N\to\infty$, so taking this limit we find
 \begin{align}
@@
 \end{align}
-where we can define $Q^2\equiv\sum_\alpha \bar s_\alpha^2$ to simplify
+where we can define the scalar $Q^2\equiv\sum_\alpha \bar s_\alpha^2$ to simplify
 \begin{align}
@@
 \end{align}
-For initial states with $\bar s_\z=0$, we thus find that
+For initial states with $\bk{\bar s_\z}=0$, we thus find that
 \begin{align}

### 2021-04-22 sun_phases/spin_model_paper/figures/make_plots.py

@@ -63,7 +63,7 @@ x_vals = 1+0.06*np.sin(5*np.pi*(y_vals-top_y)/(top_y-btm_y))
 plt.plot(x_vals, y_vals, "r", linewidth = 1)
 plt.plot([1], [btm_y], **marker_kw(dn_color))
 plt.plot([1], [top_y], **marker_kw(up_color))
-plt.annotate(r"$U_0$", (0.77, y_lim*0.9), ha = "center", va = "center")
+plt.annotate(r"$U$", (0.8, y_lim*0.9), ha = "center", va = "center")
 
 plt.axis("off")
 plt.subplots_adjust(0,0.05,1,1.2,0,0)
@@ -86,14 +86,19 @@ plt.xlim(-np.pi,np.pi)
 plt.xticks([])
 plt.yticks([])
 
-axis.annotate("", xytext = (1.1,0), xy = (1.1,1),
+ylim = axis.get_ylim()
+yrange = ylim[1] - ylim[0]
+btm = (-1-ylim[0]) / yrange
+top = (+1-ylim[0]) / yrange
+
+axis.annotate("", xytext = (1.1,btm), xy = (1.1,top),
               xycoords = "axes fraction",
               arrowprops = dict( arrowstyle = "<->" ))
-for hh in [ 0, 1 ]:
+for hh in [ btm, top ]:
     axis.annotate("", xytext = (1.05,hh), xy = (1.15,hh),
                   xycoords = "axes fraction",
                   arrowprops = dict( arrowstyle = "-" ))
-axis.annotate(r"$\sim J$", xy = (1.15,.5),
+axis.annotate(r"$4J$", xy = (1.15,.5),
               xycoords = "axes fraction",
               va = "center", ha = "left")
 
@@ -108,7 +113,7 @@ for hh in [ h1, h2 ]:
     axis.annotate("", xytext = (.45,hh), xy = (.55,hh),
                   xycoords = "axes fraction",
                   arrowprops = dict( arrowstyle = "-" ))
-axis.annotate(r"$U_0$", xy = (.62,(h1+h2)/2),
+axis.annotate(r"$U$", xy = (.62,(h1+h2)/2),
               xycoords = "axes fraction",
               va = "center", ha = "center")
 
@@ -456,14 +461,18 @@ class Arrow3D(patches.FancyArrowPatch):
         self.set_positions((xs[0],ys[0]),(xs[1],ys[1]))
[…]

### 2021-04-22 collective_sun/sun_perturbation.tex
changed some formatting and revised intro

@@
 \documentclass[nofootinbib,notitlepage,11pt]{revtex4-2}
+\usepackage[T1]{fontenc}
 
 %%% linking references
+\usepackage[dvipsnames]{xcolor}
 \usepackage{hyperref}
@@
   colorlinks=true,
-  linkcolor=blue,
-  filecolor=magenta,
-  urlcolor=cyan,
+  allcolors=BlueViolet,
 }
@@
 
-SU($n$) symmetries play an important role in physics.  Underpinning
-much of high energy physics, the celebrated SU($n$) gauge theory known
-as Yang-Mills theory is central to our understanding of the
-electroweak and strong forces.  Extensions of Yang-Mills and SU($n$)
-symmetry feature in the most well-studied examples of holographic
-duality\cite{maldacena1999largen} and the connection between
-entanglement and gravity\cite{ryu2006holographic} through the anti-de
-Sitter / conformal field theory (AdS/CFT) correspondence.  In a
-condensed matter setting, SU(2) appears ubiquitously as a symmetry of
-the Hubbard model, with important consequences for the study of
-quantum magnetism and high temperature
-superconductivity\cite{lee2006doping}.  The extension of the SU(2)
-Hubbard model to SU($n$) has led to predictions of exotic phases of
-matter such as valence bond solids\cite{read1989valencebond,
-  rokhsar1990quadratic, kaul2012lattice, hermele2011topological} and
-chiral spin liquids\cite{hermele2009mott, hermele2011topological,
-  chen2016syntheticgaugefield, nataf2016chiral}, as well as the
-potential to perform universal topological quantum
-computation\cite{freedman2004class, nayak2008nonabelian}.
-Furthermore, the consideration of disordered SU($n$) spin models has
-opened analytically tractable avenues for studying quantum chaos and
-information scrambling\cite{sachdev1993gapless,
-  bentsen2019integrable}.
-
-Last but not least, SU($n$) symmetries appear naturally in the physics
-of atomic, molecular, and optical (AMO)
-systems\cite{gorshkov2010twoorbital, beverland2016realizing,
-  cazalilla2014ultracold, taie2012su, hofrichter2016direct,
-  cappellini2014direct, scazza2014observation, zhang2014spectroscopic,
-  goban2018emergence, perlin2019effective} through the independence of
-atomic orbital and interaction parameters on the $n$ nuclear spin
-states of individual atoms, with e.g.~$n=10$ in the case of
-${}^{87}$Sr.  As a result, experiments can directly probe the SU($n$)
-Hubbard model, leading to experimental observations of SU($n$) Hubbard
-phases and phase transitions\cite{taie2012su, hofrichter2016direct},
-two-orbital SU($n$) magnetism\cite{cappellini2014direct,
-  scazza2014observation, zhang2014spectroscopic}, and multi-body
-SU($n$)-symmetric interactions\cite{goban2018emergence,
-  perlin2019effective}.  The tremendous theoretical significance of
-SU($n$) symmetries makes it all the more exciting that they appear in
-experimental platforms with exquisite degrees of microscopic control.
-This control enables detailed studies of these symmetries'
-consequences for fundamental questions in physics, as well as their
-practical use in technological applications.  SU(2)-symmetric spin
-interactions, for example, can play an important role in the
-development of quantum sensors that make use of entanglement to
-surpass classical limits on measurement
-precision\cite{he2019engineering, perlin2020spin}.
-The exciting prospect of similarly exploiting larger SU($n$)
-symmetries to achieve a technological advantage is still an unexplored
-avenue of research with enormous potential.
-
-\red{[Paragraph on collective spin systems]}
-
-In this work, we generalize some of the techniques used to harness
-SU(2)-symmetric spin interactions for metrological advantage to the
-case of SU($n$).  In doing so, we develop a low-energy effective
-theory of gapped SU($n$) ferromagnets with weakly broken SU($n$)
-symmetry (e.g.~due to external fields or additional interaction
-terms).  The restriction of our theory to ferromagnetic interactions
-can be relaxed to the assumption of
+SU($n$) symmetries play an important role in physics.
+Underpinning much of high energy physics, the SU($n$) gauge theory known as Yang-Mills theory is central to our understanding of the electroweak and strong forces.
+Extensions of Yang-Mills and SU($n$) symmetry feature in the most well-studied examples of holographic duality \cite{maldacena1999largen} and the connection between entanglement and gravity \cite{ryu2006holographic} through the anti-de Sitter/conformal field theory (AdS/CFT) correspondence.
+In a condensed matter setting, SU($2$) appears ubiquitously as a symmetry of the Hubbard model, with important consequences for the study of quantum magnetism and high temperature superconductivity \cite{lee2006doping}.
+The extension of SU($2$) Hubbard and spin models to SU($n$) has led to predictions of exotic phases of matter such as valence bond solids \cite{read1989valencebond, rokhsar1990quadratic, kaul2012lattice, hermele2011topological} and chiral spin liquids \cite{hermele2009mott, hermele2011topological, chen2016syntheticgaugefield, nataf2016chiral}, as well as the potential to perform universal topological quantum computation \cite{freedman2004class, nayak2008nonabelian} and other phenomena \cite{nataf2014exact, nataf2016exact}.
+Furthermore, disordered SU($n$) spin models have opened analytically tractable avenues for studying quantum chaos and information scrambling \cite{sachdev1993gapless, bentsen2019integrable}.
+
+The tremendous theoretical significance of SU($n$) symmetries makes it all the more exciting that they appear naturally in experimental atomic, molecular, and optical (AMO) platforms \cite{wu2003exact, cazalilla2009ultracold, gorshkov2010twoorbital, beverland2016realizing, cazalilla2014ultracold, taie2012su, hofrichter2016direct, cappellini2014direct, scazza2014observation, zhang2014spectroscopic, goban2018emergence, perlin2019effective} with exquisite degrees of microscopic control.
+As a simple example, this symmetry arises through the independence of atomic orbital and interaction parameters on the $n$ nuclear spin states of alkaline-earth(-like) atoms, with e.g.~$n=10$ for ${}^{87}$Sr \cite{wu2003exact, cazalilla2009ultracold, cazalilla2014ultracold}.
+As a result, AMO experiments can directly probe the role of SU($n$) interactions in controllable settings.
+Recent progress includes the experimental observations of bosonization and faster thermalization in multi-component Fermi gas with SU($n$) symmetry \cite{song2020evidence, sonderhouse2020thermodynamics}, SU($n$) Hubbard phases and phase transitions \cite{taie2012su, hofrichter2016direct, taie2020observation}, two-orbital SU($n$) magnetism \cite{cappellini2014direct, scazza2014observation, zhang2014spectroscopic}, and multi-body SU($n$)-symmetric interactions \cite{goban2018emergence, perlin2019effective}.
+In the spirit of quantum simulation, further investigations in controlled settings will play an important role in understanding the consequence of SU($n$) symmetries for fundamental questions in physics, as well as their practical use in technological applications.
+For example, SU($2$)-symmetric spin interactions can be harnessed to develop quantum sensors that surpass classical limits on measurement precision \cite{he2019engineering, perlin2020spin}.
+The prospect of similarly exploiting more general SU($n$) symmetries to achieve a technological advantage is still an unexplored avenue of research with untapped potential.
+
+In this work, we generalize some of the techniques used to harness SU(2)-symmetric spin interactions for metrological advantage to the case of SU($n$).
+In doing so, we develop a low-energy effective theory of gapped SU($n$) ferromagnets with weakly broken SU($n$) symmetry (e.g.~due to external fields or additional interaction terms).
+The restriction of our theory to ferromagnetic interactions can be relaxed to the assumption of
 \begin{enumerate*}
-\item an initial state in the manifold of permutationally symmetric
-  states, a multilevel generalization of the Dicke manifold for
-  two-level spins, together with the assumption that
-\item this manifold is gapped away from states that break
-  permutational symmetry.
+\item an initial state in the manifold of permutationally symmetric states, a multilevel generalization of the Dicke manifold for two-level spins, together with the assumption that
+\item this manifold is gapped away from states that break permutational symmetry.
 \end{enumerate*}
-In addition to working out the effective Hamiltonian induced by weak
-SU($n$) symmetry breaking through second order in perturbation theory,
-we develop a diagrammatic method to systematically compute matrix
-elements of arbitrary spin operators within manifolds of weakly broken
-permutational symmetry, thereby enabling the numerical simulation of
-dynamics induced by non-perturbative modifications to the SU($n$)
-Heisenberg model.  While these simulations generally accumulate
-uncontrolled errors from the truncation of Hilbert space, we identify
-criteria under which numerical results correctly capture the
-qualitative dynamical behavior of a system.
-
-In the following section, \ref{sec:bare_sun}, we introduce the SU($n$)
-Heisenberg model, finding a convenient way to express its Hamiltonian
-in terms of permutation (or SWAP) operations that are agnostic to the
-value of $n$.  In Section \ref{sec:breaking_sun}, we consider
-modifying the Hamiltonian of the bare Heisenberg model by the addition
-of spin operators, corresponding to e.g.~external fields or additional
-interactions, that weakly break SU($n$) symmetry.  We sketch out a
-derivation of the effective Hamiltonian induced by these terms through
-second order in perturbation theory, deferring more detailed
-calculations to appendices.  Finally, we discuss a numerical method to
-simulate dynamics within manifolds of weak permutational symmetry
-breaking in Section \ref{sec:shell_model}, before concluding in
-Section \ref{sec:conclusions}.  Wherever possible, we will fall back
-to the broadly familiar case of SU(2) in order to facilitate an
-understanding of the generalizations to SU($n$).
-
-\red{[re-visit / revise the introduction after finishing the remaining
-  text]}
+In addition to working out the effective Hamiltonian induced by weak SU($n$) symmetry breaking through second order in perturbation theory, we develop a diagrammatic method to systematically compute matrix elements of arbitrary spin operators within manifolds of weakly broken permutational symmetry, thereby enabling the numerical simulation of dynamics induced by non-perturbative modifications to the SU($n$) Heisenberg model.
+While these simulations generally accumulate uncontrolled errors from the truncation of Hilbert space, we identify criteria under which numerical results correctly capture the qualitative dynamical behavior of a system.
+
+In the following section, \ref{sec:bare_sun}, we introduce the SU($n$) Heisenberg model, finding a convenient way to express its Hamiltonian in terms of permutation (or SWAP) operations that are agnostic to the value of $n$.
+In Section \ref{sec:breaking_sun}, we consider modifying the Hamiltonian of the bare Heisenberg model by the addition of spin operators, corresponding to e.g.~external fields or additional interactions, that weakly break SU($n$) symmetry.
+We sketch out a derivation of the effective Hamiltonian induced by these terms through second order in perturbation theory, deferring more detailed calculations to appendices.
+Finally, we discuss a numerical method to simulate dynamics within manifolds of weak permutational symmetry breaking in Section \ref{sec:shell_model}, before concluding in Section \ref{sec:conclusions}.
+Wherever possible, we will fall back to the broadly familiar case of SU(2) in order to facilitate an understanding of the generalizations to SU($n$).
+
+\red{[re-visit / revise the introduction after finishing the remaining text]}
 
@@
 Before getting to the full SU($n$) Heisenberg model, we first revisit
-the more common case of SU(2).  The SU(2) Heisenberg model consists of
+the more common case of SU(2).
+The SU(2) Heisenberg model consists of
 $N$ two-level spins on a lattice, with spin-aligning interactions
@@
 \end{align}
-where $i,j$ index individual spins; $h_{ij}$ are scalar coefficients
-with $h_{ij}=h_{ji}$ and $h_{ii}=0$; and
-$\m s_i\equiv\p{s_\x^{(i)},s_\y^{(i)},s_\z^{(i)}}$ is a vector of spin
-operators $s_\alpha^{(i)}\equiv\sigma_\alpha^{(i)}/2$ defined by Pauli
-operators $\sigma_\alpha^{(i)}$ acting on spin $i$.  In the standard
-language of spin physics, each term of this Hamiltonian essentially
-imposes an energy penalty of $-2h_{ij}$ to the formation of a spin 0
-singlet state $\ket{\up\dn}-\ket{\dn\up}$ between spins $i$ and $j$,
-favoring the formation of spin triplet states with total spin 1.
-Another interpretation of the Hamiltonian in \eqref{eq:H_SU2} is
-illuminated by the observation that the two-spin operator
-$\m s_i\c\m s_j$ is, up to an overall constant, essentially the
-two-qubit SWAP operation $\Pi_{ij}$.  We will refer to $\Pi_{ij}$ as
-{\it permutation operator}, because its action on the Hilbert space
-$\H\equiv\bigotimes_{k=1}^N\H_k$ of $N$ spins with individual Hilbert
-spaces $\H_k$ is to permute the tensor factors of $\H$ associated with
-spins $i$ and $j$ (see Figure \ref{fig:sketch_spins}).  Up to identity
-terms with no physical consequence, we then have
+where $i,j$ index individual spins; $h_{ij}$ are scalar coefficients with $h_{ij}=h_{ji}$ and $h_{ii}=0$; and $\m s_i\equiv(s_\x^{(i)},s_\y^{(i)},s_\z^{(i)})$ is a vector of spin operators $s_\alpha^{(i)}\equiv\sigma_\alpha^{(i)}/2$ defined by Pauli operators $\sigma_\alpha^{(i)}$ acting on spin $i$.
+In the standard language of spin physics, each term of this Hamiltonian essentially imposes an energy penalty of $-2h_{ij}$ to the formation of a spin 0 singlet state $\ket{\up\dn}-\ket{\dn\up}$ between spins $i$ and $j$, favoring the formation of spin triplet states with total spin 1.
+Another interpretation of the Hamiltonian in \eqref{eq:H_SU2} is illuminated by the observation that the two-spin operator $\m s_i\c\m s_j$ is, up to an overall constant, essentially the two-qubit SWAP operation $\Pi_{ij}$.
+We will refer to $\Pi_{ij}$ as {\it permutation operator}, because its action on the Hilbert space $\H\equiv\bigotimes_{k=1}^N\H_k$ of $N$ spins with individual Hilbert spaces $\H_k$ is to permute the tensor factors of $\H$ associated with spins $i$ and $j$ (see Figure \ref{fig:sketch_spins}).
+Up to identity terms with no physical consequence, we then have
 \begin{align}
@@
 \end{align}
-where $s_{\mu\nu}^{(i)}\equiv\op{\mu}{\nu}_i$ is a spin-transition
-operator for spin $i$, and $\mu,\nu\in\set{\up,\dn}$ index the states
-of a single spin.  In addition to providing a nice interpretation of
-the Hamiltonian $H_0$ as essentially swapping the states of spins $i$
-and $j$ at a rate $h_{ij}$, the definition of the SU(2)-symmetric
-Heisenberg model in in \eqref{eq:H_0} is straightforward to generalize
-to SU($n$) with arbitrary $n$.  That is, rather than having
-$\mu,\nu\in\set{\up,\dn}$ index two spin states, we simply let
-$\mu,\nu\in\set{1,2,\cdots,n}$ index the states (sometimes called {\it
-  colors} or {\it flavors}) of an $n$-level spin.
+where $s_{\mu\nu}^{(i)}\equiv\op{\mu}{\nu}_i$ is a spin-transition operator for spin $i$, and $\mu,\nu\in\set{\up,\dn}$ index the states of a single spin.
+In addition to providing a nice interpretation of the Hamiltonian $H_0$ as essentially swapping the states of spins $i$ and $j$ at a rate $h_{ij}$, the definition of the SU(2)-symmetric Heisenberg model in in \eqref{eq:H_0} is straightforward to generalize to SU($n$) with arbitrary $n$.
+That is, rather than having $\mu,\nu\in\set{\up,\dn}$ index two spin states, we simply let $\mu,\nu\in\set{1,2,\cdots,n}$ index the states (sometimes called {\it colors} or {\it flavors}) of an $n$-level spin.
 
 \begin{figure}
-  \centering
-  \includegraphics[valign=m]{sketch_spins.pdf}
-  $\stackrel{\displaystyle\Pi_{ij}}
-  {\displaystyle\longleftrightarrow}$
-  \includegraphics[valign=m]{sketch_spins_swap.pdf}
-  \caption{The permutation operator $\Pi_{ij}$ swaps the states of
-    interacting spins $i$ and $j$, which in the case of SU(2) can be
-    represented by magnetization vectors in 3-dimensional space.  In
-    the case of SU($n$), these vectors live in a higher-dimensional
-    space without a clear geometric interpretation, but the
-    interpretation of the permutation operator $\Pi_{ij}$ remains the
-    same.}
+  \centering \includegraphics[valign=m]{sketch_spins.pdf} $\stackrel{\displaystyle\Pi_{ij}} {\displaystyle\longleftrightarrow}$ \includegraphics[valign=m]{sketch_spins_swap.pdf}
+  \caption{The permutation operator $\Pi_{ij}$ swaps the states of interacting spins $i$ and $j$, which in the case of SU(2) can be represented by magnetization vectors in 3-dimensional space.
+    In the case of SU($n$), these vectors live in a higher-dimensional space without a clear geometric interpretation, but the interpretation of the permutation operator $\Pi_{ij}$ remains the same.
+  }
 \label{fig:sketch_spins}
@@
 
-If the coefficients $h_{ij}$ of the interaction Hamiltonian $H_0$ are
-all negative, then the ground-state manifold $\M_0$ of $H_0$ consists
-of spin-polarized, {\it permutationally symmetric} (PS) states that
-are simultaneous $+1$ eigenstates of all permutation operators
-$\Pi_{ij}$.  We will assume that all $h_{ij}<0$ throughout these
-notes, but note that this restriction can be relaxed to the assumption
-that the initial state of all spins is a permutationally symmetric
-state (e.g.~a spin-polarized state).  We will also assume that the PS
-manifold $\M_0$ is gapped away from all orthogonal states by an
-interaction energy $\Delta_{\t{gap}}>0$.  Power-law couplings of the
-form $h_{ij}\sim-1/\abs{i-j}^\alpha$ on a $D$-dimensional lattice, for
-example, always yield a non-vanishing spectral gap $\Delta_{\t{gap}}$
-for long-range interactions with $\alpha\le D$ (see Appendix
-\ref{sec:spectral_gap}).
-
-In the case of SU(2), the PS manifold $\M_0$ is precisely the {\it
-  Dicke manifold} spanned by the PS states $\ket{m_\up,m_\dn}$ with a
-definite (integer) number $m_\mu\ge0$ of spins occupying the
-single-spin state $\mu\in\set{\up,\dn}$.  If there are $N$ spins
-total, then Dicke states $\ket{m_\up,m_\dn}$ must have
-$m_\up+m_\dn=N$.  In the case of multilevel spins, the PS manifold
-$\M_0$ is similarly spanned by states
-$\ket{m}=\ket{m_1,m_2,\cdots,m_n}$ with a definite occupation number
-$m_\mu\ge0$ of the single-spin state $\mu$, and $\sum_\mu m_\mu=N$.
-These states are essentially labeled by an assignment
-$m=\p{m_1,m_2,\cdots,m_n}$ of $N$ (identical) spins to $n$ (distinct)
-states.  We denote the set of all such assignments by $\A_n\p{N}$,
-such that our standard basis for $\M_0$ is
-$\set{\ket{m}:m\in\A_n\p{N}}$.  The dimension of $\M_0$, or
-equivalently the ground-state degeneracy of $H_0$, is straightforward
-to determine with a standard ``stars and bars'' calculation:
-$\dim\p{\M_0}=\abs{\A_n\p{N}} = {N+n-1 \choose n-1} \sim N^{n-1}$.
-The energy of any PS state $\ket\psi\in\M_0$ with respect to the
-interaction Hamiltonian $H_0$ in \eqref{eq:H_0} is
+If the coefficients $h_{ij}$ of the interaction Hamiltonian $H_0$ are all negative, then the ground-state manifold $\M_0$ of $H_0$ consists of spin-polarized, {\it permutationally symmetric} (PS) states that are simultaneous $+1$ eigenstates of all permutation operators $\Pi_{ij}$.
+We will assume that all $h_{ij}<0$ throughout these notes, but note that this restriction can be relaxed to the assumption that the initial state of all spins is a permutationally symmetric state (e.g.~a spin-polarized state).
+We will also assume that the PS manifold $\M_0$ is gapped away from all orthogonal states by an interaction energy $\Delta_{\t{gap}}>0$.
+Power-law couplings of the form $h_{ij}\sim-1/\abs{i-j}^\alpha$ on a $D$-dimensional lattice, for example, always yield a non-vanishing spectral gap $\Delta_{\t{gap}}$ for long-range interactions with $\alpha\le D$ (see Appendix \ref{sec:spectral_gap}).
+
+In the case of SU(2), the PS manifold $\M_0$ is precisely the {\it Dicke manifold} spanned by the PS states $\ket{m_\up,m_\dn}$ with a definite (integer) number $m_\mu\ge0$ of spins occupying the single-spin state $\mu\in\set{\up,\dn}$.
+If there are $N$ spins total, then Dicke states $\ket{m_\up,m_\dn}$ must have $m_\up+m_\dn=N$.
+In the case of multilevel spins, the PS manifold $\M_0$ is similarly spanned by states $\ket{m}=\ket{m_1,m_2,\cdots,m_n}$ with a definite occupation number $m_\mu\ge0$ of the single-spin state $\mu$, and $\sum_\mu m_\mu=N$.
+
+These states are essentially labeled by an assignment $m=\p{m_1,m_2,\cdots,m_n}$ of $N$ (identical) spins to $n$ (distinct) states.
+We denote the set of all such assignments by $\A_n\p{N}$, such that our standard basis for $\M_0$ is $\set{\ket{m}:m\in\A_n\p{N}}$.
+The dimension of $\M_0$, or equivalently the ground-state degeneracy of $H_0$, is straightforward to determine with a standard ``stars and bars'' calculation: $\dim\p{\M_0}=\abs{\A_n\p{N}} = {N+n-1 \choose n-1} \sim N^{n-1}$.
+The energy of any PS state $\ket\psi\in\M_0$ with respect to the interaction Hamiltonian $H_0$ in \eqref{eq:H_0} is
 \begin{align}
@@
 
-A common task, for example in state preparation protocols or quenched
-dynamics experiments, is to first initialize in some trivial uniform
-product state, $\ket\phi^{\otimes N}$, and then to let it evolve under
-a (possibly time-dependent) Hamiltonian that induces non-trivial
-dynamics.  The permutational symmetry of states such as
-$\ket\phi^{\otimes N}\in\M_0$, however, makes them eigenstates of the
-SU($n$)-symmetric interaction Hamiltonian $H_0$ in \eqref{eq:H_0}.  In
-order to induce non-trivial dynamics, we therefore need to add new
-terms to the Hamiltonian that break its SU($n$) symmetry.
-
-As a concrete example, this sort of scenario can be used in the SU(2)
-case to generate spin squeezed states that harness entanglement to
-surpass classical limits on measurement
-precision\cite{he2019engineering, perlin2020spin}.
-Starting with a general (gapped) SU(2)-symmetric Hamiltonian of the
-form in \eqref{eq:H_SU2} we can consider the addition of Ising
-interactions
+A common task, for example in state preparation protocols or quenched dynamics experiments, is to first initialize in some trivial uniform product state, $\ket\phi^{\otimes N}$, and then to let it evolve under a (possibly time-dependent) Hamiltonian that induces non-trivial dynamics.
+The permutational symmetry of states such as $\ket\phi^{\otimes N}\in\M_0$, however, makes them eigenstates of the SU($n$)-symmetric interaction Hamiltonian $H_0$ in \eqref{eq:H_0}.
+In order to induce non-trivial dynamics, we therefore need to add new terms to the Hamiltonian that break its SU($n$) symmetry.
+
+As a concrete example, this sort of scenario can be used in the SU(2) case to generate spin squeezed states that harness entanglement to surpass classical limits on measurement precision\cite{he2019engineering, perlin2020spin}.
+Starting with a general (gapped) SU(2)-symmetric Hamiltonian of the form in \eqref{eq:H_SU2} we can consider the addition of Ising interactions
 \begin{align}
@@
 \end{align}
-If these interactions are sufficiently weak for perturbative
-treatment, as in the case of the XXZ model near the Heisenberg point,
-then the effective Hamiltonian induced by $H_{\t{Ising}}$ on initially
-PS states $\ket\psi\in\M_0$ is the one-axis twisting (OAT)
-Hamiltonian\cite{ma2011quantum}
+If these interactions are sufficiently weak for perturbative treatment, as in the case of the XXZ model near the Heisenberg point, then the effective Hamiltonian induced by $H_{\t{Ising}}$ on initially PS states $\ket\psi\in\M_0$ is the one-axis twisting (OAT) Hamiltonian\cite{ma2011quantum}
 \begin{align}
@@
 \end{align}
-where the collective operator $S_\z\equiv\sum_{j=1}^Ns_\z^{(j)}$, and
-the squeezing strength $\chi$ is equal to the mean Ising coupling
-strength $J_{ij}$.  If Ising interactions are unavailable, the same
-effect can be achieved by applying an inhomogeneous magnetic field,
+where the collective operator $S_\z\equiv\sum_{j=1}^Ns_\z^{(j)}$, and the squeezing strength $\chi$ is equal to the mean Ising coupling strength $J_{ij}$.
+If Ising interactions are unavailable, the same effect can be achieved by applying an inhomogeneous magnetic field,
 \begin{align}
@@
 \end{align}
-which gives rise to an OAT Hamiltonian at second order in perturbation
-theory, with a squeezing strength $\chi$ that is related to the
-variance of the magnetic field strength $B_i$.
-
-Similarly to single- and two-body operators such as $H_{\t{B}}$ and
-$H_{\t{Ising}}$ in \eqref{eq:magnetic} and \eqref{eq:Ising}, it will
-be necessary for our purposes to additionally consider the
-introduction of $M$-body operators with $M>2$. Generalizing the above
-examples to the case of SU($n$) systems, we therefore consider a
-general $M$-body operator of the form
+which gives rise to an OAT Hamiltonian at second order in perturbation theory, with a squeezing strength $\chi$ that is related to the variance of the magnetic field strength $B_i$.
+
+Similarly to single- and two-body operators such as $H_{\t{B}}$ and $H_{\t{Ising}}$ in \eqref{eq:magnetic} and \eqref{eq:Ising}, it will be necessary for our purposes to additionally consider the introduction of $M$-body operators with $M>2$.
+Generalizing the above examples to the case of SU($n$) systems, we therefore consider a general $M$-body operator of the form
 \begin{align}
@@
 \end{align}
-where $\m w$ is a dimension-$M$ (i.e.~$M$-index) tensor of scalar
-coefficients $w_k\equiv w_{k_1k_2\cdots k_M}$; $X$ is an $M$-spin
-operator, e.g.~$s_\z\otimes s_\z$ in the case of Ising interactions
-with $M=2$; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is a list of the
-individual spins $k_i\in\ZZ_N$ that the operator
-$X_k\equiv X_{k_1k_2\cdots k_M}$ acts on; and
+where $\m w$ is a dimension-$M$ (i.e.~$M$-index) tensor of scalar coefficients $w_k\equiv w_{k_1k_2\cdots k_M}$; $X$ is an $M$-spin operator, e.g.~$s_\z\otimes s_\z$ in the case of Ising interactions with $M=2$; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is a list of the individual spins $k_i\in\ZZ_N$ that the operator $X_k\equiv X_{k_1k_2\cdots k_M}$ acts on; and
 \begin{align}
@@
 \end{align}
-is the strictly ``off-diagonal'' part of $\ZZ_N^M$, which is necessary
-to identify for a consistent definition of $X_k$ as an $M$-body
-operator.  In this notation, we can write the magnetic field and Ising
-Hamiltonians in \eqref{eq:magnetic} and \eqref{eq:Ising} as
-$H_{\t{B}}=\O\p{\m B,s_\z}$ and
-$H_{\t{Ising}}=\O\p{\m J,s_\z\otimes s_\z}$.
-
-If the addition $\O\p{\m w,X}$ to the SU($n$)-symmetric Hamiltonian
-$H_0$ in \eqref{eq:H_0} is sufficiently small, namely with operator
-norm $\norm{\O\p{\m w,X}}$ less than half the spectral gap
-$\Delta_{\t{gap}}$ of $H_0$, $\norm{\O\p{\m w,X}}<\Delta_{\t{gap}}/2$,
-then we can treat the effect of $\O\p{\m w,X}$ on the ground-state PS
-manifold $\M_0$ perturbatively.  The effective Hamiltonians
-$H_{\t{eff}}^{(1)}$ and $H_{\t{eff}}^{(2)}$ induced by $\O\p{\m w,X}$
-on $\M_0$ at first and second order in perturbation theory
-are\cite{bravyi2011schrieffer}
+is the strictly ``off-diagonal'' part of $\ZZ_N^M$, which is necessary to identify for a consistent definition of $X_k$ as an $M$-body operator.
+In this notation, we can write the magnetic field and Ising Hamiltonians in \eqref{eq:magnetic} and \eqref{eq:Ising} as $H_{\t{B}}=\O\p{\m B,s_\z}$ and $H_{\t{Ising}}=\O\p{\m J,s_\z\otimes s_\z}$.
+
+If the addition $\O\p{\m w,X}$ to the SU($n$)-symmetric Hamiltonian $H_0$ in \eqref{eq:H_0} is sufficiently small, namely with operator norm $\norm{\O\p{\m w,X}}$ less than half the spectral gap $\Delta_{\t{gap}}$ of $H_0$, $\norm{\O\p{\m w,X}}<\Delta_{\t{gap}}/2$, then we can treat the effect of $\O\p{\m w,X}$ on the ground-state PS manifold $\M_0$ perturbatively.
+The effective Hamiltonians $H_{\t{eff}}^{(1)}$ and $H_{\t{eff}}^{(2)}$ induced by $\O\p{\m w,X}$ on $\M_0$ at first and second order in perturbation theory are\cite{bravyi2011schrieffer}
 \begin{align}
@@
 \end{align}
-where $\P_\Delta$ is a projector onto the eigenspace $\E_\Delta$ of
-the interaction Hamiltonian $H_0$ with interaction energy $\Delta$
-above that of PS manifold $\M_0$.  In particular, $\P_0$ is a
-projector onto $\M_0$ itself, and $\E_0=\M_0$.  The first order
-effective Hamiltonian $H_{\t{eff}}^{(1)}$ simply projects
-$\O\p{\m w,X}$ onto the PS manifold $\M_0$, and takes the form
+where $\P_\Delta$ is a projector onto the eigenspace $\E_\Delta$ of the interaction Hamiltonian $H_0$ with interaction energy $\Delta$ above that of PS manifold $\M_0$.
+In particular, $\P_0$ is a projector onto $\M_0$ itself, and $\E_0=\M_0$.
+The first order effective Hamiltonian $H_{\t{eff}}^{(1)}$ simply projects $\O\p{\m w,X}$ onto the PS manifold $\M_0$, and takes the form
 \begin{align}
@@
 \end{align}
-where $\EQPS$ denotes equality under a restriction to $\M_0$; the
-coefficient $\mean w$ is the average of all $w_k$; and $\col X$ is a
-collective version of $X$:
+where $\EQPS$ denotes equality under a restriction to $\M_0$; the coefficient $\mean w$ is the average of all $w_k$; and $\col X$ is a collective version of $X$:
 \begin{align}
@@
 \end{align}
-with $\abs{\D_N\p{M}}=\prod_{j=0}^{M-1}\p{N-j}$.  In the case of a
-magnetic field $s_\z$ or Ising interactions $s_\z\otimes s_\z$, for
-example,
+with $\abs{\D_N\p{M}}=\prod_{j=0}^{M-1}\p{N-j}$.
+In the case of a magnetic field $s_\z$ or Ising interactions $s_\z\otimes s_\z$, for example,
 \begin{align}
@@
 \end{align}
-The second order effective Hamiltonian $H_{\t{eff}}^{(2)}$ in
-\eqref{eq:H_eff} takes more work to simplify due to the presence of a
-projector $\P_\Delta$ onto the manifold $\E_\Delta$ of states with
-excitation energy $\Delta$.  This projector essentially picks off the
-part of $\O\p{\m w,X}$ that is strictly off-diagonal with respect to
-the ground- and excited-state manifolds $\E_0$ and $\E_\Delta$.  We
-therefore need to decompose $\O\p{\m w,X}$ into components that
-generate states of definite excitation energy when acting on PS states
-$\ket\psi\in\M_0$.  The SU($n$) symmetry of $H_0$ enables such a
-decomposition to take the form
+The second order effective Hamiltonian $H_{\t{eff}}^{(2)}$ in \eqref{eq:H_eff} takes more work to simplify due to the presence of a projector $\P_\Delta$ onto the manifold $\E_\Delta$ of states with excitation energy $\Delta$.
+This projector essentially picks off the part of $\O\p{\m w,X}$ that is strictly off-diagonal with respect to the ground- and excited-state manifolds $\E_0$ and $\E_\Delta$.
+We therefore need to decompose $\O\p{\m w,X}$ into components that generate states of definite excitation energy when acting on PS states $\ket\psi\in\M_0$.
+The SU($n$) symmetry of $H_0$ enables such a decomposition to take the form
 \begin{align}
@@
 \end{align}
-where, thinking of the tensor $\m w$ as a $\abs{\D_N\p{M}}$-component
-vector, the tensor $\m w_\Delta$ can be found by diagonalizing a
-matrix $\hat{\m h}$ of dimensions
-$\abs{\D_N\p{M}}\times\abs{\D_N\p{M}}\sim N^M\times N^M$, and
-projecting $\m w$ onto the eigenspace of $\hat{\m h}$ with eigenvalue
-$\Delta$ (see Appendix \ref{sec:eigenstates}).  The matrix
-$\hat{\m h}$ is determined entirely by the couplings couplings
-$h_{ij}$ of $H_0$.  While the computational cost of diagonalizing
-$\hat{\m h}$ is nominally $O\p{N^{3M}}$, spatial symmetries of $H_0$
-and $\O\p{\m w,X}$ such as translational invariance or isotropy can
-drastically reduce the complexity of this eigenvalue problem, as we
-discuss in Appendix \ref{sec:symmetries}.
-
-Equipped with the decomposition of $\O\p{\m w,X}$ into terms
-$\O\p{\m w_\Delta,X}$ that generate states of definite excitation
-energy $\Delta$, we can expand
+where, thinking of the tensor $\m w$ as a $\abs{\D_N\p{M}}$-component vector, the tensor $\m w_\Delta$ can be found by diagonalizing a matrix $\hat{\m h}$ of dimensions $\abs{\D_N\p{M}}\times\abs{\D_N\p{M}}\sim N^M\times N^M$, and projecting $\m w$ onto the eigenspace of $\hat{\m h}$ with eigenvalue $\Delta$ (see Appendix \ref{sec:eigenstates}).
+The matrix $\hat{\m h}$ is determined entirely by the couplings couplings $h_{ij}$ of $H_0$.
+While the computational cost of diagonalizing $\hat{\m h}$ is nominally $O\p{N^{3M}}$, spatial symmetries of $H_0$ and $\O\p{\m w,X}$ such as translational invariance or isotropy can drastically reduce the complexity of this eigenvalue problem, as we discuss in Appendix \ref{sec:symmetries}.
+
+Equipped with the decomposition of $\O\p{\m w,X}$ into terms $\O\p{\m w_\Delta,X}$ that generate states of definite excitation energy $\Delta$, we can expand
 \begin{align}
@@
 \end{align}
-Simplifying the product $\P_0 \O\p{\m w_\Delta,X}^2 \P_0$ generally
-yields $r$-body operators for all positive integers $r\le2M$.  We
-provide a general prescription for simplifying such products in
-Appendix \ref{sec:operator_product}, and explicitly work out the
-single-body ($M=1$) and two-body ($M=2$) cases in Appendices
-\ref{sec:single_pair_prod} and \ref{sec:two_pair_prod}.  If $X$ is a
-single-body operator, then
+Simplifying the product $\P_0 \O\p{\m w_\Delta,X}^2 \P_0$ generally yields $r$-body operators for all positive integers $r\le2M$.
+We provide a general prescription for simplifying such products in Appendix \ref{sec:operator_product}, and explicitly work out the single-body ($M=1$) and two-body ($M=2$) cases in Appendices \ref{sec:single_pair_prod} and \ref{sec:two_pair_prod}.
+If $X$ is a single-body operator, then
 \begin{align}
@@
 \end{align}
-where $\mean{X\otimes X}$ is the average of
-$\p{X\otimes X}_{ij}=X_iX_j$ over all $i\ne j$, similarly to the
-definition of $\mean{w}$ in Eq.~\eqref{eq:mean_col}.  In the SU(2)
-case, this result implies that an inhomogeneous magnetic field
-($X=s_\z$) yields the OAT Hamiltonian $H_{\t{OAT}}$ in \eqref{eq:OAT}
-at second order perturbation theory.
+where $\mean{X\otimes X}$ is the average of $\p{X\otimes X}_{ij}=X_iX_j$ over all $i\ne j$, similarly to the definition of $\mean{w}$ in Eq.~\eqref{eq:mean_col}.
+In the SU(2) case, this result implies that an inhomogeneous magnetic field ($X=s_\z$) yields the OAT Hamiltonian $H_{\t{OAT}}$ in \eqref{eq:OAT} at second order perturbation theory.
 
@@
 
-In order to find the effective Hamiltonian $H_{\t{eff}}^{(2)}$ induced
-by the $M$-body operator $\O\p{\m w,X}$ on the PS manifold $\M_0$ at
-second order in perturbation theory, we constructed states with
-definite excitation energy $\Delta>0$ with respect to $H_0$.  Rather
-than treating these excited states as ``intermediate'' or ``virtual''
-for the derivation of an effective ground-state Hamiltonian, we can
-perform numerical simulations in the joint Hilbert space of both
-ground and excited states.  When the perturbation $\O\p{\m w,X}$ is
-small, namely with operator norm
-$\norm{\O\p{\m w,X}}<\Delta_{\t{gap}}/2$, such simulations are
-equivalent (up to perturbative corrections) to a higher-order
-perturbation theory.  Moreover, the first-class treatment of excited
-states in finite systems enables simulating the dynamics induced by
-{\it non-perturbative} additions $\O\p{\m w,X}$ to the base
-Hamiltonian $H_0$.
-
-The basic idea will be to identify the set of states (manifolds)
-$\M_r$ with a definite number ($r$) of spin-wave excitations, or {\it
-  magnons}.  The ground-state manifold of PS states is simply the
-zero-magnon manifold $\M_0$.  If we construct a basis for all $\M_r$
-with $r\le r_{\t{max}}$, we can perform numerical simulations in the
-combined manifold of all states with at most $r_{\t{max}}$ magnons; we
-refer to such simulations as {\it order-}$r_{\t{max}}$ {\it truncated
-  shell} (TS$_{r_{\t{max}}}$) simulations.  An $M$-spin operator
-$\O\p{\m w,X}$ can only couple $r$-magnon states (in $\M_r$) to
-$q$-magnon states (in $\M_q$) with $\abs{q-r}\le M$.  For initially PS
-states, TS$_{r_{\t{max}}}$ simulations thus capture all virtual
-processes through order $2\floor{r_{\t{max}}/M}+1$ in perturbation
-theory.  When $\norm{\O\p{\m w,X}}\ge\Delta_{\t{gap}}/2$, however,
-perturbation theory formally breaks down, and the error incurred by
-neglecting all $s$-magnon states with $s>r_{\t{max}}$ is not clear a
-priori.  Nonetheless, TS$_{r_{\t{max}}}$ simulations can be
-benchmarked by examining the populations of all $r$-magnon manifolds
-$\M_r$ with $r\le r_{\t{max}}$.  As long as the maximal populations
-within large-$r$ manifolds $\M_r$ are small, one should expect
-TS$_{r_{\t{max}}}$ simulations to faithfully capture (at least
-qualitatively) the dynamical behavior of the spin system under
-consideration.
-
-The remainder of this section is devoted to the construction of a
-basis for the $r$-magnon manifolds $\M_r$, and to the expansion of
-general spin operators $\O\p{\m w,X}$ in this basis.  We begin with a
-classification of excited states in the case of SU(2).  A common
-starting point is to first classify states by a definite
-(non-negative) spin length $S=\set{N/2,N/2-1,\cdots}$, and a spin
-projection $m_\z\in\set{S,S-1,\cdots,-S}$ onto a fixed quantization
-axis.  In fact, the manifold of states with definite fixed spin length
-$S$ is precisely the $\p{N/2-S}$-magnon manifold $\M_{N/2-S}$.  The PS
-manifold of Dicke states with $S=N/2$ is then $\M_0$, and more
-generally the manifold of $r$-magnon states with $S=N/2-r$ is $\M_r$.
-Rather than the spin projection $m_\z$, we then identify states by the
-occupation numbers $m=\p{m_\up,m_\dn}\in\A_2\p{N}$, where $m_\mu$
-denotes the total number of spins in the state $\mu$, with
-$m_\up+m_\dn=N$ and $m_\up-m_\dn=m_\z$.
-
-The label $m$ suffices to identify the PS (Dicke) manifold
-$\M_0=\oper{span}\set{\ket{m}:m\in\A_2\p{N}}$, but we need additional
-labels to uniquely identify states in the $r$-magnon manifold $\M_r$
-with $r>0$.  Single-magnon ($r=1$) spin-wave states of SU(2) spin
-systems can be written in the form
+In order to find the effective Hamiltonian $H_{\t{eff}}^{(2)}$ induced by the $M$-body operator $\O\p{\m w,X}$ on the PS manifold $\M_0$ at second order in perturbation theory, we constructed states with definite excitation energy $\Delta>0$ with respect to $H_0$.
+Rather than treating these excited states as ``intermediate'' or ``virtual'' for the derivation of an effective ground-state Hamiltonian, we can perform numerical simulations in the joint Hilbert space of both ground and excited states.
+When the perturbation $\O\p{\m w,X}$ is small, namely with operator norm $\norm{\O\p{\m w,X}}<\Delta_{\t{gap}}/2$, such simulations are equivalent (up to perturbative corrections) to a higher-order perturbation theory.
+Moreover, the first-class treatment of excited states in finite systems enables simulating the dynamics induced by {\it non-perturbative} additions $\O\p{\m w,X}$ to the base Hamiltonian $H_0$.
+
+The basic idea will be to identify the set of states (manifolds) $\M_r$ with a definite number ($r$) of spin-wave excitations, or {\it magnons}.
+The ground-state manifold of PS states is simply the zero-magnon manifold $\M_0$.
+If we construct a basis for all $\M_r$ with $r\le r_{\t{max}}$, we can perform numerical simulations in the combined manifold of all states with at most $r_{\t{max}}$ magnons; we refer to such simulations as {\it order-}$r_{\t{max}}$ {\it truncated shell} (TS$_{r_{\t{max}}}$) simulations.
+An $M$-spin operator $\O\p{\m w,X}$ can only couple $r$-magnon states (in $\M_r$) to $q$-magnon states (in $\M_q$) with $\abs{q-r}\le M$.
+For initially PS states, TS$_{r_{\t{max}}}$ simulations thus capture all virtual processes through order $2\floor{r_{\t{max}}/M}+1$ in perturbation theory.
+When $\norm{\O\p{\m w,X}}\ge\Delta_{\t{gap}}/2$, however, perturbation theory formally breaks down, and the error incurred by neglecting all $s$-magnon states with $s>r_{\t{max}}$ is not clear a priori.
+Nonetheless, TS$_{r_{\t{max}}}$ simulations can be benchmarked by examining the populations of all $r$-magnon manifolds $\M_r$ with $r\le r_{\t{max}}$.
+As long as the maximal populations within large-$r$ manifolds $\M_r$ are small, one should expect TS$_{r_{\t{max}}}$ simulations to faithfully capture (at least qualitatively) the dynamical behavior of the spin system under consideration.
+
+The remainder of this section is devoted to the construction of a basis for the $r$-magnon manifolds $\M_r$, and to the expansion of general spin operators $\O\p{\m w,X}$ in this basis.
+We begin with a classification of excited states in the case of SU(2).
+A common starting point is to first classify states by a definite (non-negative) spin length $S=\set{N/2,N/2-1,\cdots}$, and a spin projection $m_\z\in\set{S,S-1,\cdots,-S}$ onto a fixed quantization axis.
+In fact, the manifold of states with definite fixed spin length $S$ is precisely the $\p{N/2-S}$-magnon manifold $\M_{N/2-S}$.
+The PS manifold of Dicke states with $S=N/2$ is then $\M_0$, and more generally the manifold of $r$-magnon states with $S=N/2-r$ is $\M_r$.
+Rather than the spin projection $m_\z$, we then identify states by the occupation numbers $m=\p{m_\up,m_\dn}\in\A_2\p{N}$, where $m_\mu$ denotes the total number of spins in the state $\mu$, with $m_\up+m_\dn=N$ and $m_\up-m_\dn=m_\z$.
+
+The label $m$ suffices to identify the PS (Dicke) manifold $\M_0=\oper{span}\set{\ket{m}:m\in\A_2\p{N}}$, but we need additional labels to uniquely identify states in the $r$-magnon manifold $\M_r$ with $r>0$.
+Single-magnon ($r=1$) spin-wave states of SU(2) spin systems can be written in the form
 \begin{align}
@@
 where each $v_j$ is a scalar and $\sigma_\z$ is a Pauli-$z$ operator.
-If an SU(2) spin system is translationally invariant, then spin-wave
-eigenstates have $v_j=e^{\i k_{\m v}\c j}$ with wavenumber
-$k_{\m v}\ne0$.  Similarly, $r$-magnon states for SU(2) take the form
-$\ket{m,\m w,\sigma_\z^{\otimes r}} \equiv \O\p{\m
-  w,\sigma_\z^{\otimes r}}\ket{m}$, where now $\m w$ is an $r$-index
-tensor that encodes the structure of correlations in the excited state
-$\ket{m,\m w,\sigma_\z^{\otimes r}}$.  The coefficient tensor $\m w$
-determines the energy of the excited state
-$\ket{m,\m w,\sigma_\z^{\otimes r}}$ through the expansion in
-\eqref{eq:excitation}.  Considering states of this form, built using
-tensor products of Pauli-$z$ ($\sigma_\z$) operators, suffices to
-construct a basis for all excited states of an SU(2) spin system (see
-Appendix \ref{sec:operator_content}).  Note that $r$-body operators
-with arbitrary coefficients $\m w$ will generally couple the PS
-manifold $\M_0$ to all $q$-magnon manifolds $\M_q$ with $q\le r$ (see
-Figure \ref{fig:manifold_sketch}).  In order to generalize the notion
-of $r$-magnon manifolds $\M_r$ to the case of SU($n$) with $n>2$,
-where the analogous notion of a spin length $S$ is less intuitive, we
-can define $\M_r$ for $r>0$ as the manifold of states that (i) can be
-generated by the action of $r$-body operators on $\M_0$, and (ii) are
-orthogonal to all states in $\bigoplus_{q<r}\M_q$.
+If an SU(2) spin system is translationally invariant, then spin-wave eigenstates have $v_j=e^{\i k_{\m v}\c j}$ with wavenumber $k_{\m v}\ne0$.
+Similarly, $r$-magnon states for SU(2) take the form $\ket{m,\m w,\sigma_\z^{\otimes r}} \equiv \O\p{\m w,\sigma_\z^{\otimes r}}\ket{m}$, where now $\m w$ is an $r$-index tensor that encodes the structure of correlations in the excited state $\ket{m,\m w,\sigma_\z^{\otimes r}}$.
+The coefficient tensor $\m w$ determines the energy of the excited state $\ket{m,\m w,\sigma_\z^{\otimes r}}$ through the expansion in \eqref{eq:excitation}.
+Considering states of this form, built using tensor products of Pauli-$z$ ($\sigma_\z$) operators, suffices to construct a basis for all excited states of an SU(2) spin system (see Appendix \ref{sec:operator_content}).
+Note that $r$-body operators with arbitrary coefficients $\m w$ will generally couple the PS manifold $\M_0$ to all $q$-magnon manifolds $\M_q$ with $q\le r$ (see Figure \ref{fig:manifold_sketch}).
+In order to generalize the notion of $r$-magnon manifolds $\M_r$ to the case of SU($n$) with $n>2$, where the analogous notion of a spin length $S$ is less intuitive, we can define $\M_r$ for $r>0$ as the manifold of states that (i) can be generated by the action of $r$-body operators on $\M_0$, and (ii) are orthogonal to all states in $\bigoplus_{q<r}\M_q$.
 
@@
   \end{tikzpicture}
-  \caption{Schematic of relationships between $q$-magnon manifolds
-    $\M_q$ with total spin $S=N/2-q$ in an SU(2) spin system.
-    Single-body operators $\O\p{\m v,\sigma_\z}$ generally couple the
-    ground-state manifold $\M_0$ to the single-magnon manifold $\M_1$
-    as well as $\M_0$ itself, while $r$-body operators
-    $\O\p{\m w,\sigma_\z^{\otimes r}}$ generally couple $\M_0$ to all
-    $q$-magnon manifolds $\M_q$ with $q\le r$.}
+  \caption{Schematic of relationships between $q$-magnon manifolds $\M_q$ with total spin $S=N/2-q$ in an SU(2) spin system.
+    Single-body operators $\O\p{\m v,\sigma_\z}$ generally couple the ground-state manifold $\M_0$ to the single-magnon manifold $\M_1$ as well as $\M_0$ itself, while $r$-body operators $\O\p{\m w,\sigma_\z^{\otimes r}}$ generally couple $\M_0$ to all $q$-magnon manifolds $\M_q$ with $q\le r$.}
   \label{fig:manifold_sketch}
@@
 
-In the case of an SU($n$) spin system, we can similarly start with the
-ground-state PS manifold
-$\M_0=\oper{span}\set{\ket{m}:m\in\A_n\p{N}}$, and generate excited
-states by the action of multi-body operators,
-$\ket{m,\m w,X}\equiv\O\p{\m w,X}\ket{m}$.  The choice of tensors
-$\m w$ for generating states in the $r$-magnon manifold $\M_r$ is the
-same as before, but we now have the additional complication of having
-to choose $r$-spin operators, $X$.  As with the choice of
-$\sigma_\z^{\otimes r}$ for SU(2), it suffices to choose tensor
-products of {\it diagonal} single-spin operators (that is, diagonal in
-a fixed single-particle basis), but there will now be $\sim\p{n-1}^r$
-choices: one of $n-1$ diagonal operators (that are linearly
-independent of the identity) for each of $r$ tensor factors.  We
-discuss the exact choice of operators in more detail in Appendix
-\ref{sec:operator_content}.  To summarize, we define $n$ single-spin
-operators $Z_\nu\equiv1-2\op\nu=\sum_{\mu\ne\nu}\op\mu-\op\nu$, for
-which $Z_\nu=\pm\sigma_\z$ in the case of SU(2), and we choose a list
-of non-negative integers $z\equiv\p{z_1,z_2,\cdots,z_n}$ with
-$\sum_\nu z_\nu=r$ to define the $r$-spin operator
+In the case of an SU($n$) spin system, we can similarly start with the ground-state PS manifold $\M_0=\oper{span}\set{\ket{m}:m\in\A_n\p{N}}$, and generate excited states by the action of multi-body operators, $\ket{m,\m w,X}\equiv\O\p{\m w,X}\ket{m}$.
+The choice of tensors $\m w$ for generating states in the $r$-magnon manifold $\M_r$ is the same as before, but we now have the additional complication of having to choose $r$-spin operators, $X$.
+As with the choice of $\sigma_\z^{\otimes r}$ for SU(2), it suffices to choose tensor products of {\it diagonal} single-spin operators (that is, diagonal in a fixed single-particle basis), but there will now be $\sim\p{n-1}^r$ choices: one of $n-1$ diagonal operators (that are linearly independent of the identity) for each of $r$ tensor factors.
+We discuss the exact choice of operators in more detail in Appendix \ref{sec:operator_content}.
+To summarize, we define $n$ single-spin operators $Z_\nu\equiv1-2\op\nu=\sum_{\mu\ne\nu}\op\mu-\op\nu$, for which $Z_\nu=\pm\sigma_\z$ in the case of SU(2), and we choose a list of non-negative integers $z\equiv\p{z_1,z_2,\cdots,z_n}$ with $\sum_\nu z_\nu=r$ to define the $r$-spin operator
 \begin{align}
@@
 \end{align}
-Our starting basis for $\M_r$ then consists of states
-$\ket{m,\m w,X_z}$, where we enforce that $z_\nu\le m_\nu$ for all
-$\nu$ (where $m_\nu$ is the occupation number of the single-spin state
-$\nu$), and that $z$ is zero in an entry where $m$ is maximal:
-$z_\nu=0$ for $\nu=\min\set{\mu:m_\mu=\max\p{m}}$.  This second
-restriction is equivalent to removing an operator from
-$\set{Z_\nu:\nu\in\ZZ_n}$, making the remaining set linearly
-independent of the identity operator.  The basis
-$\set{\ket{m,\m w,X_z}}$ for $\M_r$ is generally complete, but not
-orthonormal, so as a final step we use the states $\ket{m,\m w,X_z}$
-to (numerically) construct an orthonormal basis $\B_r$ for $\M_r$ (see
-Appendix \ref{sec:operator_content}).
-
-By repeating the above process incrementally for all
-$r\le r_{\t{max}}$ (starting with $r=1$), we can construct an
-orthonormal basis
-$\col\B_{r_{\t{max}}}\equiv\bigoplus_{r\le r_{\t{max}}}\B_r$ for the
-combined manifold of all states with at most $r_{\t{max}}$ magnons.
-During the construction of $\col\B_s$, an appropriate choice of
-tensors $\m w$ ensures that all states $\ket{m,\m w,X}\in\col\B_s$ are
-eigenstates of the base Hamiltonian $H_0$ (see Appendix
-\ref{sec:eigenstates}), which is to say that $H_0$ is diagonal with
-respect to $\col\B_s$.  Matrix elements of general spin operators
-$\O\p{\m v,Y}$, meanwhile, take the form
+Our starting basis for $\M_r$ then consists of states $\ket{m,\m w,X_z}$, where we enforce that $z_\nu\le m_\nu$ for all $\nu$ (where $m_\nu$ is the occupation number of the single-spin state $\nu$), and that $z$ is zero in an entry where $m$ is maximal: $z_\nu=0$ for $\nu=\min\set{\mu:m_\mu=\max\p{m}}$.
+This second restriction is equivalent to removing an operator from $\set{Z_\nu:\nu\in\ZZ_n}$, making the remaining set linearly independent of the identity operator.
+The basis $\set{\ket{m,\m w,X_z}}$ for $\M_r$ is generally complete, but not orthonormal, so as a final step we use the states $\ket{m,\m w,X_z}$ to (numerically) construct an orthonormal basis $\B_r$ for $\M_r$ (see Appendix \ref{sec:operator_content}).
+
+By repeating the above process incrementally for all $r\le r_{\t{max}}$ (starting with $r=1$), we can construct an orthonormal basis $\col\B_{r_{\t{max}}}\equiv\bigoplus_{r\le r_{\t{max}}}\B_r$ for the combined manifold of all states with at most $r_{\t{max}}$ magnons.
+During the construction of $\col\B_s$, an appropriate choice of tensors $\m w$ ensures that all states $\ket{m,\m w,X}\in\col\B_s$ are eigenstates of the base Hamiltonian $H_0$ (see Appendix \ref{sec:eigenstates}), which is to say that $H_0$ is diagonal with respect to $\col\B_s$.
+Matrix elements of general spin operators $\O\p{\m v,Y}$, meanwhile, take the form
 \begin{align}
@@
 \end{align}
-In order to compute these matrix elements, we simplify general
-products of multi-body operators in Appendix
-\ref{sec:operator_product}.
+In order to compute these matrix elements, we simplify general products of multi-body operators in Appendix \ref{sec:operator_product}.
 
@@
 \begin{acknowledgments}
-  We thank Asier Pi\~neiro Orioli for helpful discussions and
-  technical contributions to this work.
+  We thank Asier Pi\~neiro Orioli for helpful discussions and technical contributions to this work.
 \end{acknowledgments}
@@
 
-Here we show that in $D$ spatial dimensions, power-law
-SU($n$)-symmetric interactions of the form in \eqref{eq:H_0} with
-$h_{ij}=-1/\abs{i-j}^\alpha$ yield a non-vanishing spectral gap when
-$\alpha\le D$.  For definiteness, we will consider a periodic
-hypercubic lattice of $N=L^D$ spins.  The spectral gap of the
-interaction Hamiltonian is then determined by well-known spin-wave
-theory (see Ref.~[\citenum{ashcroft1978solid}] or Appendix
-\ref{sec:single_body_eigenstates})
+Here we show that in $D$ spatial dimensions, power-law SU($n$)-symmetric interactions of the form in \eqref{eq:H_0} with $h_{ij}=-1/\abs{i-j}^\alpha$ yield a non-vanishing spectral gap when $\alpha\le D$.
+For definiteness, we will consider a periodic hypercubic lattice of $N=L^D$ spins.
+The spectral gap of the interaction Hamiltonian is then determined by well-known spin-wave theory (see Ref.~[\citenum{ashcroft1978solid}] or Appendix \ref{sec:single_body_eigenstates})
 \begin{align}
@@
 \end{align}
-where $\ZZ_L$ denotes the set of integers modulo $L$;
-$k_{\t{SE}}\in\ZZ_L^D\times2\pi/L$ is a wavenumber for a single-magnon
-spin-wave state; and the norm $\abs{d}$ for a vector $d$ on a periodic
-lattice is implicitly understood to mean the smallest Euclidean
-distance of $d$ from a fixed origin.  In order to yield the smallest
-excitation energy $\Delta$, the wavenumber $k_{\t{SE}}$ should
-maximize the contribution of the cosine term above, which is achieved
-by a wavenumber that minimizes the oscillations of this term when
-integrated over the entire lattice.  A suitable candidate for a
-minimal wavenumber is $k_{\t{SE}}=\p{2\pi/L,0,0,\cdots}$, which
-corresponds to an excitation energy
+where $\ZZ_L$ denotes the set of integers modulo $L$; $k_{\t{SE}}\in\ZZ_L^D\times2\pi/L$ is a wavenumber for a single-magnon spin-wave state; and the norm $\abs{d}$ for a vector $d$ on a periodic lattice is implicitly understood to mean the smallest Euclidean distance of $d$ from a fixed origin.
+In order to yield the smallest excitation energy $\Delta$, the wavenumber $k_{\t{SE}}$ should maximize the contribution of the cosine term above, which is achieved by a wavenumber that minimizes the oscillations of this term when integrated over the entire lattice.
+A suitable candidate for a minimal wavenumber is $k_{\t{SE}}=\p{2\pi/L,0,0,\cdots}$, which corresponds to an excitation energy
 \begin{align}
@@
 \end{align}
-Defining $\epsilon\equiv2/L$ and a rescaled domain symmetrized about
-$0$, $\SS_\epsilon\simeq\ZZ_L/\epsilon$ with
-$\SS_\epsilon\subset\sp{-1,1}$, we substitute $x\simeq\epsilon d$ to
-get
+Defining $\epsilon\equiv2/L$ and a rescaled domain symmetrized about $0$, $\SS_\epsilon\simeq\ZZ_L/\epsilon$ with $\SS_\epsilon\subset\sp{-1,1}$, we substitute $x\simeq\epsilon d$ to get
 \begin{align}
@@
 \end{align}
-As $\epsilon\to0$ the discrete sum over $x$ is well approximated by an
-integral that avoids an infinitesimal region about the origin, i.e.
+As $\epsilon\to0$ the discrete sum over $x$ is well approximated by an integral that avoids an infinitesimal region about the origin, i.e.
 \begin{align}
@@
 \end{align}
-where the integral $\I_D\p{\epsilon}$ is defined using the symmetric
-interval $\TT_a\equiv\p{-a/2,a/2}$.  The integrand of
-$\I_D\p{\epsilon}$ is strictly positive and well-behaved on the
-entirety of its domain except for the origin, where depending on the
-value of $\alpha$ the integrand may vanish or diverge as
-$\abs{x}\to0$.  Together, these facts mean that
+where the integral $\I_D\p{\epsilon}$ is defined using the symmetric interval $\TT_a\equiv\p{-a/2,a/2}$.
+The integrand of $\I_D\p{\epsilon}$ is strictly positive and well-behaved on the entirety of its domain except for the origin, where depending on the value of $\alpha$ the integrand may vanish or diverge as $\abs{x}\to0$.
+Together, these facts mean that
 \begin{align}
@@
 \end{align}
-for some $\gamma\ge0$, which implies that gap $\Delta$ is
-non-vanishing when $\alpha\le D\le D+\gamma$.
-
-For the sake of completion, we add that in fact the gap $\Delta$
-always vanishes in the thermodynamic limit when $\alpha>D$.  To see
-this behavior, we note that the asymptotic dependence of the integral
-$\I_D\p{\epsilon}$ on $\epsilon$ is determined by the behavior of its
-integrand when $\abs{x}\sim\epsilon$, in which case
-$1-\cos\p{\pi x_1}\sim x_1^2$, so
+for some $\gamma\ge0$, which implies that gap $\Delta$ is non-vanishing when $\alpha\le D\le D+\gamma$.
+
+For the sake of completion, we add that in fact the gap $\Delta$ always vanishes in the thermodynamic limit when $\alpha>D$.
+To see this behavior, we note that the asymptotic dependence of the integral $\I_D\p{\epsilon}$ on $\epsilon$ is determined by the behavior of its integrand when $\abs{x}\sim\epsilon$, in which case $1-\cos\p{\pi x_1}\sim x_1^2$, so
 \begin{align}
@@
 \end{align}
-We can then use the fact that $x_1^2\le\abs{x}^2$ and change to
-spherical coordinates to find that
+We can then use the fact that $x_1^2\le\abs{x}^2$ and change to spherical coordinates to find that
 \begin{align}
@@
 
-Both perturbation theory and beyond-perturbative simulations of the
-dynamics induced by an operator $\O\p{\m w,X}$ require us to decompose
-$\O\p{\m w,X}$ into a sum of terms that generate eigenstates of the
-SU($n$)-symmetric interaction Hamiltonian $H_0$ in \eqref{eq:H_0} when
-acting on permutationally symmetric (PS) states $\ket\psi\in\M_0$.  To
-perform this decomposition, in this section we simplify products of
-the form $H_0\O\p{\m w,X}\ket\psi$ with $\ket\psi\in\M_0$.  For
-reference, the $M$-body operator $\O\p{\m w,X}$ is defined by
+Both perturbation theory and beyond-perturbative simulations of the dynamics induced by an operator $\O\p{\m w,X}$ require us to decompose $\O\p{\m w,X}$ into a sum of terms that generate eigenstates of the SU($n$)-symmetric interaction Hamiltonian $H_0$ in \eqref{eq:H_0} when acting on permutationally symmetric (PS) states $\ket\psi\in\M_0$.
+To perform this decomposition, in this section we simplify products of the form $H_0\O\p{\m w,X}\ket\psi$ with $\ket\psi\in\M_0$.
+For reference, the $M$-body operator $\O\p{\m w,X}$ is defined by
 \begin{align}
@@
 \end{align}
-where $\m w$ is a dimension-$M$ (i.e.~$M$-index) tensor of scalar
-coefficients $w_k\equiv w_{k_1k_2\cdots k_M}$; $X$ is an $M$-spin
-operator; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is a list of the individual
-spins $k_i\in\ZZ_N$ that the operator $X_k\equiv X_{k_1k_2\cdots k_M}$
-acts on; and
+where $\m w$ is a dimension-$M$ (i.e.~$M$-index) tensor of scalar coefficients $w_k\equiv w_{k_1k_2\cdots k_M}$; $X$ is an $M$-spin operator; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is a list of the individual spins $k_i\in\ZZ_N$ that the operator $X_k\equiv X_{k_1k_2\cdots k_M}$ acts on; and
 \begin{align}
@@
 
-To build familiarity with the problem at hand, we first consider the
-simple case of a single-body ($M=1$) case and simplify
+To build familiarity with the problem at hand, we first consider the simple case of a single-body ($M=1$) case and simplify
 \begin{align}
@@
 \end{align}
-The sum in \eqref{eq:single_body_diagnosis_start} has terms with
-$k\in\set{i,j}$, and terms with $k\notin\set{i,j}$.  In the case of
-$k\notin\set{i,j}$, the permutation operator $\Pi_{ij}$ commutes with
-$X_k$ and annihilates on $\ket\psi$, and we can replace the sum
+The sum in \eqref{eq:single_body_diagnosis_start} has terms with $k\in\set{i,j}$, and terms with $k\notin\set{i,j}$.
+In the case of $k\notin\set{i,j}$, the permutation operator $\Pi_{ij}$ commutes with $X_k$ and annihilates on $\ket\psi$, and we can replace the sum
 \begin{align}
@@
 \end{align}
-where $E_0=\frac12\sum_{i\ne j}h_{ij}$ is the interaction energy of PS
-states $\ket\psi\in\M_0$.  Making the replacement
+where $E_0=\frac12\sum_{i\ne j}h_{ij}$ is the interaction energy of PS states $\ket\psi\in\M_0$.
+Making the replacement
 \begin{align}
@@
 \end{align}
-which implies that the terms in \eqref{eq:single_body_diagnosis_start}
-with $k\notin\set{i,j}$ are
+which implies that the terms in \eqref{eq:single_body_diagnosis_start} with $k\notin\set{i,j}$ are
 \begin{align}
@@
 \end{align}
-The terms in \eqref{eq:single_body_diagnosis_start} with
-$k\in\set{i,j}$, meanwhile, are
+The terms in \eqref{eq:single_body_diagnosis_start} with $k\in\set{i,j}$, meanwhile, are
 \begin{align}
@@
 \end{align}
-The action of the single-body perturbation $\O\p{\m w,X}$ on a
-permutationally symmetric state therefore generates an eigenstate of
-$H_0$ with energy $E_0+\Delta$ if the vector $\m w=\sum_k w_k \ket{k}$
-satisfies the eigenvalue equation
+The action of the single-body perturbation $\O\p{\m w,X}$ on a permutationally symmetric state therefore generates an eigenstate of $H_0$ with energy $E_0+\Delta$ if the vector $\m w=\sum_k w_k \ket{k}$ satisfies the eigenvalue equation
 \begin{align}
@@
 \end{align}
-where $\m h\equiv\sum_{i,j} h_{ij} \op{i}{j}$ is a matrix of all
-couplings $h_{ij}$; the vector
-$\v h \equiv \sum_{i,j} h_{ij} \ket{i} = \sum_i h_i \ket{i}$ is the
-sum of all columns of $\m h$; and $\oper{diag}\v h$ is a matrix with $\v h$
-on the diagonal and zeroes everywhere else.  When the interaction
-Hamiltonian $H_0$ is translationally invariant, the single-body
-eigenvalue problem in \eqref{eq:single_body_eig} is solvable
-analytically.  In this case, the couplings $h_{ij}$ depend only on the
-separation $\abs{i-j}$, so eigenvectors of $\m h$ are plane waves of
-the form
+where $\m h\equiv\sum_{i,j} h_{ij} \op{i}{j}$ is a matrix of all couplings $h_{ij}$; the vector $\v h \equiv \sum_{i,j} h_{ij} \ket{i} = \sum_i h_i \ket{i}$ is the sum of all columns of $\m h$; and $\oper{diag}\v h$ is a matrix with $\v h$ on the diagonal and zeroes everywhere else.
+When the interaction Hamiltonian $H_0$ is translationally invariant, the single-body eigenvalue problem in \eqref{eq:single_body_eig} is solvable analytically.
+In this case, the couplings $h_{ij}$ depend only on the separation $\abs{i-j}$, so eigenvectors of $\m h$ are plane waves of the form
 \begin{align}
@@
 \end{align}
-where on a $D$-dimensional periodic lattice of $N=L^D$ spins, lattice
-sites are indexed by vectors $d\in\ZZ_L^D$, and wavenumbers take on
-values $k\in\ZZ_L^D\times2\pi/L$.  The corresponding eigenvalues of
-$\m h$ can be determined by expanding
+where on a $D$-dimensional periodic lattice of $N=L^D$ spins, lattice sites are indexed by vectors $d\in\ZZ_L^D$, and wavenumbers take on values $k\in\ZZ_L^D\times2\pi/L$.
+The corresponding eigenvalues of $\m h$ can be determined by expanding
 \begin{align}
@@
 \end{align}
-where the imaginary contributions vanish in the sum over $d$ because
-$h_{0,d}=h_{0,-d}$.  The remainder of \eqref{eq:single_body_eig} that
-we need to sort out is $\oper{diag}\v h$, where all
-$h_i = \sum_{i,j}h_{ij} = \sum_d h_{0,d}$ are equal, which implies
-that $\oper{diag}\v h = \sum_d h_{0,d}$ is a scalar.  We thus find that
+where the imaginary contributions vanish in the sum over $d$ because $h_{0,d}=h_{0,-d}$.
+The remainder of \eqref{eq:single_body_eig} that we need to sort out is $\oper{diag}\v h$, where all $h_i = \sum_{i,j}h_{ij} = \sum_d h_{0,d}$ are equal, which implies that $\oper{diag}\v h = \sum_d h_{0,d}$ is a scalar.
+We thus find that
 \begin{align}
@@
 as the set of all subsets (``choices'') of $q$ elements from $\ZZ_N$.
-The sum in \eqref{eq:diagnosis_start} has terms with $i,j\notin k$,
-terms with $i,j\in k$, and terms with only one of $i$ or $j\in k$.
-The permutation operator $\Pi_{ij}$ acts trivially on $X_k\ket\psi$
-when $i,j\notin k$, and we can replace the sum
+The sum in \eqref{eq:diagnosis_start} has terms with $i,j\notin k$, terms with $i,j\in k$, and terms with only one of $i$ or $j\in k$.
+The permutation operator $\Pi_{ij}$ acts trivially on $X_k\ket\psi$ when $i,j\notin k$, and we can replace the sum
 \begin{align}
@@
 \end{align}
-where $E_0 = \sum_{\set{i,j}\in\C_N\p{2}} h_{ij}$ is the interaction
-energy of PS states $\ket\psi\in\M_0$, and
+where $E_0 = \sum_{\set{i,j}\in\C_N\p{2}} h_{ij}$ is the interaction energy of PS states $\ket\psi\in\M_0$, and
 \begin{align}
@@
 \end{align}
-with $k_a,k_b\in k\equiv\p{k_1,k_2,\cdots,k_M}$.  Defining a linear
-operator $\hat{\m h}_{\t{D}}$ on the space of dimension-$M$ tensors,
-whose action on $\m w$ is a new dimension-$M$ tensor with entries
+with $k_a,k_b\in k\equiv\p{k_1,k_2,\cdots,k_M}$.
+Defining a linear operator $\hat{\m h}_{\t{D}}$ on the space of dimension-$M$ tensors, whose action on $\m w$ is a new dimension-$M$ tensor with entries
 \begin{align}
@@
 \end{align}
-we can write the terms in \eqref{eq:diagnosis_start} with
-$i,j\notin k$ as
+we can write the terms in \eqref{eq:diagnosis_start} with $i,j\notin k$ as
 \begin{align}
@@
 \end{align}
-where $k_{a\lra b}$ denotes a list that is equal to
-$k=\p{k_1,k_2,\cdots,k_M}$ but with $k_a$ and $k_b$ swapped, and we
-have relabeled $k_a$ and $k_b$ to change $w_k X_{k_{a\lra b}}$ to
-$w_{k_{a\lra b}} X_k$.  To write down these terms succinctly, we
-define a linear operator $\hat{\m h}_{\t{P}}$ on the space of
-dimension-$M$ tensors, whose action on $\m w$ is a new dimension-$M$
-tensor with entries
+where $k_{a\lra b}$ denotes a list that is equal to $k=\p{k_1,k_2,\cdots,k_M}$ but with $k_a$ and $k_b$ swapped, and we have relabeled $k_a$ and $k_b$ to change $w_k X_{k_{a\lra b}}$ to $w_{k_{a\lra b}} X_k$.
+To write down these terms succinctly, we define a linear operator $\hat{\m h}_{\t{P}}$ on the space of dimension-$M$ tensors, whose action on $\m w$ is a new dimension-$M$ tensor with entries
 \begin{align}
@@
 \end{align}
-Finally, the terms in \eqref{eq:diagnosis_start} with only one of
-$i\in k$ or $j\in k$ are
+Finally, the terms in \eqref{eq:diagnosis_start} with only one of $i\in k$ or $j\in k$ are
 \begin{align}
@@
 \end{align}
-where $k_{a:i}$ denotes a list that is equal to $k$ but with $k_a$
-replaced by $i$, and we have relabeled $k_a$ and $i$ to change
-$w_k X_{k_{a:i}}$ to $w_{k_{a:i}} X_k$.  For each fixed $a$, the sum
-over $i$ above is essentially a contraction of the matrix $\m h$ with
-the $a$-th index of the tensor $\m w$.  Defining a linear operator
-$\hat{\m h}_a$ on the space of dimension-$M$ tensors, whose action on
-$\m w$ is a new dimension-$M$ tensor with entries
+where $k_{a:i}$ denotes a list that is equal to $k$ but with $k_a$ replaced by $i$, and we have relabeled $k_a$ and $i$ to change $w_k X_{k_{a:i}}$ to $w_{k_{a:i}} X_k$.
+For each fixed $a$, the sum over $i$ above is essentially a contraction of the matrix $\m h$ with the $a$-th index of the tensor $\m w$.
+Defining a linear operator $\hat{\m h}_a$ on the space of dimension-$M$ tensors, whose action on $\m w$ is a new dimension-$M$ tensor with entries
 \begin{align}
@@
 \end{align}
-where $\m w_\Delta$ is the projection of $\m w$ onto the eigenspace of
-$\hat{\m h}$ with eigenvalue $\Delta$, i.e.
+where $\m w_\Delta$ is the projection of $\m w$ onto the eigenspace of $\hat{\m h}$ with eigenvalue $\Delta$, i.e.
 \begin{align}
@@
 \end{align}
-For reference, here $k_{a:i}$ is equal to
-$k\equiv\p{k_1,k_2,\cdots,k_M}$ but with $k_a$ replaced by $i$;
-$\C_M\p{2}$ is defined in \eqref{eq:choices}; $k_{a\lra b}$ is equal
-to $k$ but with $k_a$ and $k_b$ swapped; and $h^{\t{D}}_k$ is defined
-in \eqref{eq:multi_body_op_diag_element}.
-
-If the coefficient tensor $\m w$ is permutationally symmetric, meaning
-that $\m w$ is invariant under arbitrary permutations of its tensor
-factors, then this symmetry is preserved by $\hat{\m h}$.  In this
-case, we can replace sums over $k\in\D_N\p{M}$ in
-\eqref{eq:multi_body_problem} and \eqref{eq:multi_body_mat} by sums
-over $k\in\C_N\p{M}$, and replace vectors
-$\ket{k_1,k_2,\cdots,k_M}\to\ket{\set{k_1,k_2,\cdots,k_M}}$, such that
-e.g.~$\ket{k_{a\lra b}}=\ket{k}$.  These replacements reduce the size
-of $\hat{\m h}$ from $\abs{\D_N\p{M}}\times\abs{\D_N\p{M}}$ to
-$\abs{\C_N\p{M}}\times\abs{\C_N\p{M}}$, where
-$\abs{\D_N\p{M}}=\prod_{j=0}^{M-1}\p{N-j}=M!\times{N\choose M}$ and
-$\abs{\C_N\p{M}}={N\choose M}$.  We discuss further reduction of the
-eigenvalue problem in \eqref{eq:multi_body_problem} by use of
-additional symmetries in Section \ref{sec:symmetries}.
-
-Some special cases of practical interest: if $M=1$, then
-$h^{\t{D}}_k=h_k$, so
+For reference, here $k_{a:i}$ is equal to $k\equiv\p{k_1,k_2,\cdots,k_M}$ but with $k_a$ replaced by $i$; $\C_M\p{2}$ is defined in \eqref{eq:choices}; $k_{a\lra b}$ is equal to $k$ but with $k_a$ and $k_b$ swapped; and $h^{\t{D}}_k$ is defined in \eqref{eq:multi_body_op_diag_element}.
+
+If the coefficient tensor $\m w$ is permutationally symmetric, meaning that $\m w$ is invariant under arbitrary permutations of its tensor factors, then this symmetry is preserved by $\hat{\m h}$.
+In this case, we can replace sums over $k\in\D_N\p{M}$ in \eqref{eq:multi_body_problem} and \eqref{eq:multi_body_mat} by sums over $k\in\C_N\p{M}$, and replace vectors $\ket{k_1,k_2,\cdots,k_M}\to\ket{\set{k_1,k_2,\cdots,k_M}}$, such that e.g.~$\ket{k_{a\lra b}}=\ket{k}$.
+These replacements reduce the size of $\hat{\m h}$ from $\abs{\D_N\p{M}}\times\abs{\D_N\p{M}}$ to $\abs{\C_N\p{M}}\times\abs{\C_N\p{M}}$, where $\abs{\D_N\p{M}}=\prod_{j=0}^{M-1}\p{N-j}=M!\times{N\choose M}$ and $\abs{\C_N\p{M}}={N\choose M}$.
+We discuss further reduction of the eigenvalue problem in \eqref{eq:multi_body_problem} by use of additional symmetries in Section \ref{sec:symmetries}.
+
+Some special cases of practical interest: if $M=1$, then $h^{\t{D}}_k=h_k$, so
 \begin{align}
@@
 \end{align}
-which is precisely what we found in \eqref{eq:single_body_eig}.  If
-$M=2$, then $h_{k\ell}^{\t{D}}=h_k+h_\ell-h_{k\ell}$, so
+which is precisely what we found in \eqref{eq:single_body_eig}.
+If $M=2$, then $h_{k\ell}^{\t{D}}=h_k+h_\ell-h_{k\ell}$, so
 \begin{align}
@@
 
-Solving the eigenvalue problem in \eqref{eq:multi_body_problem} to
-find $M$-body operators $\O\p{\m w_\Delta,X}$ that generate states of
-definite interaction energy nominally requires diagonalizing the
-matrix $\hat{\m h}$ with dimensions $\sim N^M\times N^M$.  Spatial
-symmetries of the interaction Hamiltonian $H_0$, however, generally
-endow $\hat{\m h}$ with a block-diagonal structure that can be used to
-reduce the complexity of this eigenvalue problem.  Translational
-invariance, for example, can reduce this problem to that of
-diagonalizing a matrix with dimensions $\sim N^{M-1}\times N^{M-1}$,
-and isotropy can further reduce the problem size by factors that
-depend on the lattice geometry.  The complexity of diagonalizing a
-matrix with dimensions $d\times d$ is $O\p{d^3}$, so even a reduction
-of $d$ by a factor of 2 yields practically significant gains (a factor
-of $2^3=8$) in the runtime of diagonalization.  Here, we provide a
-general prescription for exploiting the spatial symmetries of $H_0$ to
-simplify the eigenvalue problem in \eqref{eq:multi_body_problem}.
-
-Our task is essentially to express the multi-body eigenvalue problem
-in \eqref{eq:multi_body_problem} in a basis that respects the spatial
-symmetries of $H_0$.  These symmetries essentially break the set
-$\D_N\p{M}$ into equivalence classes $\S_M$\footnote{For brevity, we
-  will suppress the explicit dependence of $\S_M$ on $N$, as well as
-  other factors such as lattice geometry.}, where elements $s\in\S_M$
-are subsets $s\subset\D_N\p{M}$ that are invariant under the
-appropriate symmetry group action.  In the case of $M=2$ in a
-translationally invariant and isotropic system, for example, each
-equivalence class $s\in\S_2$ would consist of all pairs
-$\p{i,j}\in\D_N\p{2}$ that are a fixed distance $\abs{i-j}=d_s$ apart,
-i.e.~$s=\set{\p{i,j}\in\D_N\p{2}:\abs{i-j}=d_s}$.  If the tensor
-$\m w$ is invariant under arbitrary permutations of its tensor
-factors, this symmetry is preserved by $\hat{\m h}$, and we can place
-all $k\in\D_N\p{M}$ that are equal up to permutation into the same
-equivalence class.
-
-For simplicity, we first assume that the $M$-body operator
-$\O\p{\m w,X}$ obeys the same spatial symmetries as $H_0$.  For any
-equivalence class $s\in\S_M$, we then define a uniform superposition
-$\ket{s}$ over all $k\in s$, as well as a projector $\P_\S$ onto the
-space of uniform superpositions:
+Solving the eigenvalue problem in \eqref{eq:multi_body_problem} to find $M$-body operators $\O\p{\m w_\Delta,X}$ that generate states of definite interaction energy nominally requires diagonalizing the matrix $\hat{\m h}$ with dimensions $\sim N^M\times N^M$.
+Spatial symmetries of the interaction Hamiltonian $H_0$, however, generally endow $\hat{\m h}$ with a block-diagonal structure that can be used to reduce the complexity of this eigenvalue problem.
+Translational invariance, for example, can reduce this problem to that of diagonalizing a matrix with dimensions $\sim N^{M-1}\times N^{M-1}$, and isotropy can further reduce the problem size by factors that depend on the lattice geometry.
+The complexity of diagonalizing a matrix with dimensions $d\times d$ is $O\p{d^3}$, so even a reduction of $d$ by a factor of 2 yields practically significant gains (a factor of $2^3=8$) in the runtime of diagonalization.
+Here, we provide a general prescription for exploiting the spatial symmetries of $H_0$ to simplify the eigenvalue problem in \eqref{eq:multi_body_problem}.
+
+Our task is essentially to express the multi-body eigenvalue problem in \eqref{eq:multi_body_problem} in a basis that respects the spatial symmetries of $H_0$.
+These symmetries essentially break the set $\D_N\p{M}$ into equivalence classes $\S_M$\footnote{For brevity, we will suppress the explicit dependence of $\S_M$ on $N$, as well as other factors such as lattice geometry.}, where elements $s\in\S_M$ are subsets $s\subset\D_N\p{M}$ that are invariant under the appropriate symmetry group action.
+In the case of $M=2$ in a translationally invariant and isotropic system, for example, each equivalence class $s\in\S_2$ would consist of all pairs $\p{i,j}\in\D_N\p{2}$ that are a fixed distance $\abs{i-j}=d_s$ apart, i.e.~$s=\set{\p{i,j}\in\D_N\p{2}:\abs{i-j}=d_s}$.
+If the tensor $\m w$ is invariant under arbitrary permutations of its tensor factors, this symmetry is preserved by $\hat{\m h}$, and we can place all $k\in\D_N\p{M}$ that are equal up to permutation into the same equivalence class.
+
+For simplicity, we first assume that the $M$-body operator $\O\p{\m w,X}$ obeys the same spatial symmetries as $H_0$.
+For any equivalence class $s\in\S_M$, we then define a uniform superposition $\ket{s}$ over all $k\in s$, as well as a projector $\P_\S$ onto the space of uniform superpositions:
 \begin{align}
@@
 \end{align}
-where $\abs{s}$ is the number of elements in $s$.  In turn, we expand
-a vector of $M$-body operator coefficients
-$\m w=\sum_{k\in\D_N\p{M}}w_k\ket{k}$ in this symmetrized basis:
+where $\abs{s}$ is the number of elements in $s$.
+In turn, we expand a vector of $M$-body operator coefficients $\m w=\sum_{k\in\D_N\p{M}}w_k\ket{k}$ in this symmetrized basis:
 \begin{align}
@@
 \end{align}
-where all $w_k$ with $k\in s$ are equal by assumption, so
-$w_s\equiv\sum_{k\in s}w_k/\abs{s}=w_\ell$ for any $\ell\in s$.
-Finally, we expand the projection of $\hat{\m h}$ onto the symmetric
-subspace:
+where all $w_k$ with $k\in s$ are equal by assumption, so $w_s\equiv\sum_{k\in s}w_k/\abs{s}=w_\ell$ for any $\ell\in s$.
+Finally, we expand the projection of $\hat{\m h}$ onto the symmetric subspace:
 \begin{align}
@@
 \end{align}
-where $h^{\t{D}}_s\equiv h^{\t{D}}_k$ for any $k\in s$, with
-$h^{\t{D}}_k$ defined in \eqref{eq:multi_body_op_diag_element}; and
-$\sp{k}$ denotes the equivalence class of $k$.  The reduced multi-body
-eigenvalue problem is then
+where $h^{\t{D}}_s\equiv h^{\t{D}}_k$ for any $k\in s$, with $h^{\t{D}}_k$ defined in \eqref{eq:multi_body_op_diag_element}; and $\sp{k}$ denotes the equivalence class of $k$.
+The reduced multi-body eigenvalue problem is then
 \begin{align}
@@
 \end{align}
-More generally, even if the couplings $\m w$ of $\O\p{\m w,X}$ do not
-obey the spatial symmetries as $H_0$, these symmetries still endow
-$\hat{\m h}$ with a block-diagonal structure that reduces the
-complexity of the multi-body eigenvalue problem.  We leave this
-reduction to future work.  % todo: perform reduction
+More generally, even if the couplings $\m w$ of $\O\p{\m w,X}$ do not obey the spatial symmetries as $H_0$, these symmetries still endow $\hat{\m h}$ with a block-diagonal structure that reduces the complexity of the multi-body eigenvalue problem.
+We leave this reduction to future work.
+% todo: perform reduction
 
@@
 
-Throughout our construction of excitation energy eigenstates of the
-form $\O\p{\m w,X}\ket\psi$, where $\ket\psi\in\M_0$ is a PS state, we
-have thus far primarily considered the choice of scalar coefficients,
-$\m w$.  Here, we discuss the choice of operators, $X$, which allows
-us to construct a complete and convenient basis for the set of all
-excited states.  Our construction of a basis for the set of excited
-states begins with the standard basis $\set{\ket{m}:m\in\A_n\p{N}}$
-for the PS manifold $\M_0$, where $m\equiv\p{m_1,m_2,\cdots,m_n}$ and
-$m_\mu$ is the (integer) occupation number of the single-spin state
-$\mu$.  Written in the ``computational basis'' of $N$ multilevel
-spins, these states are
+Throughout our construction of excitation energy eigenstates of the form $\O\p{\m w,X}\ket\psi$, where $\ket\psi\in\M_0$ is a PS state, we have thus far primarily considered the choice of scalar coefficients, $\m w$.
+Here, we discuss the choice of operators, $X$, which allows us to construct a complete and convenient basis for the set of all excited states.
+Our construction of a basis for the set of excited states begins with the standard basis $\set{\ket{m}:m\in\A_n\p{N}}$ for the PS manifold $\M_0$, where $m\equiv\p{m_1,m_2,\cdots,m_n}$ and $m_\mu$ is the (integer) occupation number of the single-spin state $\mu$.
+Written in the ``computational basis'' of $N$ multilevel spins, these states are
 \begin{align}
@@
 \end{align}
-where $\C\p{m}$ is a normalization factor equal to the number of terms
-in the sum.  While $\ket{m}$ is the only PS state with occupation
-numbers $m$, the set of {\it all} states with occupation numbers $m$
-is spanned by states of the form
+where $\C\p{m}$ is a normalization factor equal to the number of terms in the sum.
+While $\ket{m}$ is the only PS state with occupation numbers $m$, the set of {\it all} states with occupation numbers $m$ is spanned by states of the form
 \begin{align}
@@
 \end{align}
-where $s\p{\Pi}\in\set{\pm1}$ for all permutations $\Pi$.  As a
-consequence, when generating excited states of the form
-$\O\p{\m w,X}\ket{m}$, we can freely restrict ourselves to operators
-$X$ that only change the signs of individual terms (associated with
-permutations $\Pi$) in \eqref{eq:pops_expansion}, which suffices to
-construct any state $\ket{m,s}$ in \eqref{eq:pops_states}.  We
-therefore consider single-operators $X$ that have $\pm1$ on the
-diagonal (in the computational basis), and zero everywhere else.
-
-The set of all operators that are $\pm1$ on the diagonal is spanned by
-sums and tensor products of the single-spin operators
-$Z_\nu\equiv1-2\op\nu=\sum_{\mu\ne\nu}\op\mu-\op\nu$.  Given a list of
-non-negative integers $z\equiv\p{z_1,z_2,\cdots,z_n}$ with
-$\sum_\mu z_\mu=r$, we thus define the $r$-spin operator
+where $s\p{\Pi}\in\set{\pm1}$ for all permutations $\Pi$.
+As a consequence, when generating excited states of the form $\O\p{\m w,X}\ket{m}$, we can freely restrict ourselves to operators $X$ that only change the signs of individual terms (associated with permutations $\Pi$) in \eqref{eq:pops_expansion}, which suffices to construct any state $\ket{m,s}$ in \eqref{eq:pops_states}.
+We therefore consider single-operators $X$ that have $\pm1$ on the diagonal (in the computational basis), and zero everywhere else.
+
+The set of all operators that are $\pm1$ on the diagonal is spanned by sums and tensor products of the single-spin operators $Z_\nu\equiv1-2\op\nu=\sum_{\mu\ne\nu}\op\mu-\op\nu$.
+Given a list of non-negative integers $z\equiv\p{z_1,z_2,\cdots,z_n}$ with $\sum_\mu z_\mu=r$, we thus define the $r$-spin operator
 \begin{align}
@@
 \end{align}
-and nominally use states of the form
-$\ket{m,\m w,z}=\O\p{\m w,X_z}\ket{m}$ as our basis for the $r$-magnon
-manifold $\M_r$.  There are three points to address concerning the
-states $\ket{m,\m w,z}$:
+and nominally use states of the form $\ket{m,\m w,z}=\O\p{\m w,X_z}\ket{m}$ as our basis for the $r$-magnon manifold $\M_r$.
+There are three points to address concerning the states $\ket{m,\m w,z}$:
 \begin{enumerate*}
-\item depending on the value of $m$, some choices of $z$ result in
-  trivial states, $\ket{m,\m w,z}=0$;
-  \label{pt:trivial}
-\item the set of nontrivial states is generally overcomplete, in that
-  any given state can be written as a nontrivial linear combination of
-  others, and finally
-  \label{pt:dependent}
-\item even after identifying a linearly independent set of nontrivial
-  states, any given pair of states $\ket{m,\m w,y},\ket{m,\m w,z}$
-  with $y\ne z$ generally have a non-zero overlap, which is to say
-  that they are not orthogonal.
-  \label{pt:orthogonal}
+\item \label{pt:trivial} depending on the value of $m$, some choices of $z$ result in trivial states, $\ket{m,\m w,z}=0$;
+\item \label{pt:dependent} the set of nontrivial states is generally overcomplete, in that any given state can be written as a nontrivial linear combination of others, and finally
+\item \label{pt:orthogonal} even after identifying a linearly independent set of nontrivial states, any given pair of states $\ket{m,\m w,y},\ket{m,\m w,z}$ with $y\ne z$ generally have a non-zero overlap, which is to say that they are not orthogonal.
 \end{enumerate*}
-To address the first two points, \ref{pt:trivial} and
-\ref{pt:dependent}, we will place restrictions on the values of $z$
-that we allow for any given $m$.  The last point, \ref{pt:orthogonal},
-will require us to construct a basis of orthonormal states numerically
-for any given values of $m,\m w$.
-
-First, we identify criteria under which we are guaranteed that
-$\ket{m,\m w,z}=0$.  We begin by assuming that $\ket{m,\m w,z}\in\M_r$
-with $\sum_\nu z_\nu=r$.  We then observe that if $z_\nu>m_\nu$, then
-$m_\nu-z_\nu$ factors of $Z_\nu$ in $X_z$ are ``wasted,'' in the sense
-that they never yield relative signs in different terms of $\ket{m}$
-when expressed in the computational basis, as in
-\eqref{eq:pops_expansion}.  More concretely, consider the individual
-terms of the expansion
+To address the first two points, \ref{pt:trivial} and \ref{pt:dependent}, we will place restrictions on the values of $z$ that we allow for any given $m$.
+The last point, \ref{pt:orthogonal}, will require us to construct a basis of orthonormal states numerically for any given values of $m,\m w$.
+
+First, we identify criteria under which we are guaranteed that $\ket{m,\m w,z}=0$.
+We begin by assuming that $\ket{m,\m w,z}\in\M_r$ with $\sum_\nu z_\nu=r$.
+We then observe that if $z_\nu>m_\nu$, then $m_\nu-z_\nu$ factors of $Z_\nu$ in $X_z$ are ``wasted,'' in the sense that they never yield relative signs in different terms of $\ket{m}$ when expressed in the computational basis, as in \eqref{eq:pops_expansion}.
+More concretely, consider the individual terms of the expansion
 \begin{align}
@@
 \end{align}
-where $\sp{X_z}_k$ denotes the action of $X_z$ on spins
-$k=\p{k_1,k_2,\cdots,k_r}$.  All terms in
-\eqref{eq:excited_state_expansion} have exactly $m_\nu$ spins in the
-state $\nu$.  For any fixed $k$, at least $m_\nu-z_\nu$ factors of
-$Z_\nu$ in $\sp{X_z}_k$ must therefore evaluate to $+1$.  We can
-eliminate these ``extra'' factors of $Z_\nu$ by replacing
-$m_\nu-z_\nu$ of the factors that evaluate to $+1$ by the identity
-operator.  Performing this replacement for all $\nu\in\ZZ_n$ yields a
-state $\ket{m,\m w_{mz},{\floor{z}_m}}=\ket{m,\m w,z}$, where
-$\floor{z}_m\equiv\p{\min\set{z_\nu,m_\nu}:\nu\in\ZZ_n}$ indexes the
-$q$-spin operator $X_{\floor{z}_m}$, and $\m w_{mz}$ is some new
-tensor of coefficients.  If $q<r$ (because $z_\nu>m_\nu$ for some
-$\nu$), then our initial assumption that $\ket{m,\m w,z}\in\M_r$
-implies that $\ket{m,\m w,z}=0$.  In other words, a nontrivial basis
-$\set{\ket{m,\m w,z}}$ for $\M_r$ must have $z_\nu\le m_\nu$ for all
-$\nu$.
-
-Next, we address the fact that the basis $\set{\ket{m,\m w,z}}$ for
-$\M_r$ is overcomplete.  This fact is a simple consequence of our
-choice of single-spin operators for the construction of
-$\ket{m,\m w,z}$, namely $\set{Z_\nu:\nu\in\ZZ_n}$.  In essence, our
-problem is the fact that $Z_\nu=\p{n-2}-\sum_{\mu\ne\nu}Z_\mu$, which
-implies that any state $\ket{m,\m w,z}\in\M_r$ can be written as a
-linear combination of states in which $z_\nu\to z_\nu-1$, and possibly
-$z_\mu\to z_\mu+1$ for some $\mu\ne\nu$.  We can eliminate this
-redundancy by removing an operator from the $\set{Z_\nu}$, or
-equivalently enforcing that $z_\nu=0$ for some $\nu$.  As a matter of
-choice, we therefore consider only states $\ket{m,\m w,z}$ for which
-$z$ is zero at the first entry where $m$ is maximal: $z_\nu=0$ for
-$\nu=\min\set{\mu:m_\mu=\max\p{m}}$.  Given our previous restriction
-that $z_\nu\le m_\nu$ for all $\nu$, enforcing that $z$ is zero at an
-entry where $m$ is maximal maximizes the number of candidate values of
-$z$ that we eliminate from consideration.
-
-While the above restrictions on $z$ eliminate some trivial and
-redundant states from the basis $\set{\ket{m,\m w,z}}$ for $\M_r$,
-they do not guarantee that there will be {\it no} trivial states or
-redundancies left over.  Furthermore, even after restrictions the
-remaining basis is not orthogonal, which is to say that generally
-$\bk{m,\m v,y|m,\m w,z}\ne0$ even when $y\ne z$.  In order to
-construct an orthonormal basis for $\M_r$, we therefore need to build
-a matrix $B_m$ with matrix elements
+where $\sp{X_z}_k$ denotes the action of $X_z$ on spins $k=\p{k_1,k_2,\cdots,k_r}$.
+All terms in \eqref{eq:excited_state_expansion} have exactly $m_\nu$ spins in the state $\nu$.
+For any fixed $k$, at least $m_\nu-z_\nu$ factors of $Z_\nu$ in $\sp{X_z}_k$ must therefore evaluate to $+1$.
+We can eliminate these ``extra'' factors of $Z_\nu$ by replacing $m_\nu-z_\nu$ of the factors that evaluate to $+1$ by the identity operator.
+Performing this replacement for all $\nu\in\ZZ_n$ yields a state $\ket{m,\m w_{mz},{\floor{z}_m}}=\ket{m,\m w,z}$, where $\floor{z}_m\equiv\p{\min\set{z_\nu,m_\nu}:\nu\in\ZZ_n}$ indexes the $q$-spin operator $X_{\floor{z}_m}$, and $\m w_{mz}$ is some new tensor of coefficients.
+If $q<r$ (because $z_\nu>m_\nu$ for some $\nu$), then our initial assumption that $\ket{m,\m w,z}\in\M_r$ implies that $\ket{m,\m w,z}=0$.
+In other words, a nontrivial basis $\set{\ket{m,\m w,z}}$ for $\M_r$ must have $z_\nu\le m_\nu$ for all $\nu$.
+
+Next, we address the fact that the basis $\set{\ket{m,\m w,z}}$ for $\M_r$ is overcomplete.
+This fact is a simple consequence of our choice of single-spin operators for the construction of $\ket{m,\m w,z}$, namely $\set{Z_\nu:\nu\in\ZZ_n}$.
+In essence, our problem is the fact that $Z_\nu=\p{n-2}-\sum_{\mu\ne\nu}Z_\mu$, which implies that any state $\ket{m,\m w,z}\in\M_r$ can be written as a linear combination of states in which $z_\nu\to z_\nu-1$, and possibly $z_\mu\to z_\mu+1$ for some $\mu\ne\nu$.
+We can eliminate this redundancy by removing an operator from the $\set{Z_\nu}$, or equivalently enforcing that $z_\nu=0$ for some $\nu$.
+As a matter of choice, we therefore consider only states $\ket{m,\m w,z}$ for which $z$ is zero at the first entry where $m$ is maximal: $z_\nu=0$ for $\nu=\min\set{\mu:m_\mu=\max\p{m}}$.
+Given our previous restriction that $z_\nu\le m_\nu$ for all $\nu$, enforcing that $z$ is zero at an entry where $m$ is maximal maximizes the number of candidate values of $z$ that we eliminate from consideration.
+
+While the above restrictions on $z$ eliminate some trivial and redundant states from the basis $\set{\ket{m,\m w,z}}$ for $\M_r$, they do not guarantee that there will be {\it no} trivial states or redundancies left over.
+Furthermore, even after restrictions the remaining basis is not orthogonal, which is to say that generally $\bk{m,\m v,y|m,\m w,z}\ne0$ even when $y\ne z$.
+In order to construct an orthonormal basis for $\M_r$, we therefore need to build a matrix $B_m$ with matrix elements
 \begin{align}
@@
 \end{align}
-where $y,z$ satisfy appropriate restrictions, and $\m v,\m w$ are
-chosen from the set of tensors that generate eigenstates of $H_0$.
-Eigenvectors of $B_m$ are then mutually orthogonal vectors of the form
-$\sum_{\m w,z}c_{\m w z}\ket{m,\m w,z}$, whose squared norms are equal
-to their corresponding eigenvalues.  Note that the matrix $B_m$ has a
-block-diagonal structure, as it cannot (for example) couple states
-$\ket{m,\m v,y},\ket{m,\m w,z}$ whose tensors $\m v,\m w$ generate
-states with different interaction energies.  We therefore only need to
-build and diagonalize this matrix independently within each block.
+where $y,z$ satisfy appropriate restrictions, and $\m v,\m w$ are chosen from the set of tensors that generate eigenstates of $H_0$.
+Eigenvectors of $B_m$ are then mutually orthogonal vectors of the form $\sum_{\m w,z}c_{\m w z}\ket{m,\m w,z}$, whose squared norms are equal to their corresponding eigenvalues.
+Note that the matrix $B_m$ has a block-diagonal structure, as it cannot (for example) couple states $\ket{m,\m v,y},\ket{m,\m w,z}$ whose tensors $\m v,\m w$ generate states with different interaction energies.
+We therefore only need to build and diagonalize this matrix independently within each block.
 
@@
 
-The construction of a basis $\B_r$ for the $r$-magnon manifold $\M_r$
-requires multiplying two multi-body operators
-$\O\p{\m v,X_y}^\dag, \O\p{\m w,X_z}$, and projecting the result onto
-the permutationally symmetric (PS) manifold $\M_0$, as in
-\eqref{eq:basis_matrix}.  Similarly, computing matrix elements of
-operators $\O\p{\m v,Y}$ with respect to the basis
-$\col\B_{r_{\t{max}}}\equiv\bigotimes_{r\le r_{\t{max}}}\B_r$ requires
-multiplying three multi-body operators and projecting the result onto
-$\M_0$, as in \eqref{eq:matrix_element}.  In this section, we develop
-a diagrammatic technique for the sorts of calculations, and work out
-some special cases of common interest.
+The construction of a basis $\B_r$ for the $r$-magnon manifold $\M_r$ requires multiplying two multi-body operators $\O\p{\m v,X_y}^\dag, \O\p{\m w,X_z}$, and projecting the result onto the permutationally symmetric (PS) manifold $\M_0$, as in \eqref{eq:basis_matrix}.
+Similarly, computing matrix elements of operators $\O\p{\m v,Y}$ with respect to the basis $\col\B_{r_{\t{max}}}\equiv\bigotimes_{r\le r_{\t{max}}}\B_r$ requires multiplying three multi-body operators and projecting the result onto $\M_0$, as in \eqref{eq:matrix_element}.
+In this section, we develop a diagrammatic technique for the sorts of calculations, and work out some special cases of common interest.
 
@@
 
-Given a system of $N$ spins, we wish to project a product of
-multi-body operators onto the PS manifold.  For $p$ multi-body
-operators, such a product takes the form
+Given a system of $N$ spins, we wish to project a product of multi-body operators onto the PS manifold.
+For $p$ multi-body operators, such a product takes the form
 \begin{align}
@@
 \end{align}
-where $\m w\equiv\p{w_1,w_2,\cdots,w_p}$ is a vector of
-dimension-$M_\alpha$ tensors $w_\alpha$;
-$\m X\equiv\p{X_1,X_2,\cdots,X_p}$ is a vector of $M_\alpha$-spin
-operators $X_\alpha$;
-$k_\alpha=\p{k_{\alpha,1},k_{\alpha,2},\cdots,k_{\alpha,M_\alpha}}$ is
-a choice of $M_\alpha$ spins that $X_\alpha\p{k_\alpha}$ acts on;
-$w_\alpha\p{k_\alpha}$ is a scalar entry of $w_\alpha$; and
-$\D_N\p{M_\alpha}$ is the strictly off-diagonal part of $\ZZ_N^M$,
-defined in \eqref{eq:off_diags}.  For later convenience, we extend the
-definitions of $w_\alpha\p{k_\alpha}=X_\alpha\p{k_\alpha}$ to all
-$k_\alpha\in\ZZ_N^M$ by defining
-$w_\alpha\p{k_\alpha}=X_\alpha\p{k_\alpha}=0$ for any $k_\alpha$ with
-repeated entries.
-
-To simplify $\Q\p{\m w,\m X}$, we first collect terms to express
-\eqref{eq:sym_prod_start} as a sum of products:
+where $\m w\equiv\p{w_1,w_2,\cdots,w_p}$ is a vector of dimension-$M_\alpha$ tensors $w_\alpha$; $\m X\equiv\p{X_1,X_2,\cdots,X_p}$ is a vector of $M_\alpha$-spin operators $X_\alpha$; $k_\alpha=\p{k_{\alpha,1},k_{\alpha,2},\cdots,k_{\alpha,M_\alpha}}$ is a choice of $M_\alpha$ spins that $X_\alpha\p{k_\alpha}$ acts on; $w_\alpha\p{k_\alpha}$ is a scalar entry of $w_\alpha$; and $\D_N\p{M_\alpha}$ is the strictly off-diagonal part of $\ZZ_N^M$, defined in \eqref{eq:off_diags}.
+For later convenience, we extend the definitions of $w_\alpha\p{k_\alpha}=X_\alpha\p{k_\alpha}$ to all $k_\alpha\in\ZZ_N^M$ by defining $w_\alpha\p{k_\alpha}=X_\alpha\p{k_\alpha}=0$ for any $k_\alpha$ with repeated entries.
+
+To simplify $\Q\p{\m w,\m X}$, we first collect terms to express \eqref{eq:sym_prod_start} as a sum of products:
 \begin{align}
@@
 \end{align}
-where $\m M\equiv\p{M_1,M_2,\cdots,M_p}$ is a vector of the dimensions
-$M_\alpha$; and
-$\D_N\p{\m M}\equiv\bigotimes_{M_\alpha\in\m M}\D_N\p{M_\alpha}$, such
-that each $k\in\D_N\p{\m M}$ has the decomposition
-$k=\p{k_1,k_2,\cdots,k_p}$ with $k_\alpha\in\D_N\p{M_\alpha}$.  We
-then classify the terms in \eqref{eq:sym_prod_sum} by the numbers
-$g_S$ of indices shared by all tensors $w_\alpha$ with
-$\alpha\in S\subset\ZZ_p$.  For example, a term of the form
-$w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}$ with distinct indices
-$a,b,c,d,e$ would have
+where $\m M\equiv\p{M_1,M_2,\cdots,M_p}$ is a vector of the dimensions $M_\alpha$; and $\D_N\p{\m M}\equiv\bigotimes_{M_\alpha\in\m M}\D_N\p{M_\alpha}$, such that each $k\in\D_N\p{\m M}$ has the decomposition $k=\p{k_1,k_2,\cdots,k_p}$ with $k_\alpha\in\D_N\p{M_\alpha}$.
+We then classify the terms in \eqref{eq:sym_prod_sum} by the numbers $g_S$ of indices shared by all tensors $w_\alpha$ with $\alpha\in S\subset\ZZ_p$.
+For example, a term of the form $w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}$ with distinct indices $a,b,c,d,e$ would have
 \begin{align}
@@
 \end{align}
-and $g_S=0$ for all other subsets $S\subset\ZZ_3$.  This index
-assignment can be associated with the Venn diagram
+and $g_S=0$ for all other subsets $S\subset\ZZ_3$.
+This index assignment can be associated with the Venn diagram
 \begin{align}
@@
 \end{align}
-where $g_S$ is determined the number of dots at the intersection of
-circles $\alpha\in S$.  We denote the set of all nonempty subsets of
-$\ZZ_p$ by $\PP^*\p{\ZZ_p}$, and denote a choice of $g_S$ for all
-$S\in\PP^*\p{\ZZ_p}$ by $g$.  With representations such as
-\eqref{eq:venn_diagram} in mind, we will loosely refer to $g$ as
-simply a ``diagram''.  We define
-$\abs{g}\equiv\sum_{S\in\PP^*\p{\ZZ_p}}g_S$ to be the number of
-indices considered in $g$ (or equivalently the number of dots in the
-diagram $g$), and we denote the set of all diagrams for
-$p=\dim\p{\m M}$ tensors with dimensions $\m M$ by $\G\p{\m M}$.  For
-consistency, all diagrams $g\in\G\p{\m M}$ must associate exactly
-$M_\alpha$ indices with the dimension-$M_\alpha$ tensor $w_\alpha$,
-i.e.
+where $g_S$ is determined the number of dots at the intersection of circles $\alpha\in S$.
+We denote the set of all nonempty subsets of $\ZZ_p$ by $\PP^*\p{\ZZ_p}$, and denote a choice of $g_S$ for all $S\in\PP^*\p{\ZZ_p}$ by $g$.
+With representations such as \eqref{eq:venn_diagram} in mind, we will loosely refer to $g$ as simply a ``diagram''.
+We define $\abs{g}\equiv\sum_{S\in\PP^*\p{\ZZ_p}}g_S$ to be the number of indices considered in $g$ (or equivalently the number of dots in the diagram $g$), and we denote the set of all diagrams for $p=\dim\p{\m M}$ tensors with dimensions $\m M$ by $\G\p{\m M}$.
+For consistency, all diagrams $g\in\G\p{\m M}$ must associate exactly $M_\alpha$ indices with the dimension-$M_\alpha$ tensor $w_\alpha$, i.e.
 \begin{align}
@@
 \end{align}
-for all $\alpha\in\ZZ_p$.  A classification of the terms in
-\eqref{eq:sym_prod_sum} by diagrams $g\in\G\p{\m M}$ allows us to
-expand
+for all $\alpha\in\ZZ_p$.
+A classification of the terms in \eqref{eq:sym_prod_sum} by diagrams $g\in\G\p{\m M}$ allows us to expand
 \begin{align}
@@
 \end{align}
-where $\EQPS$ denotes equality under a restriction to $\M_0$, and
-$\Q_g\p{\m w,\m X}$ is the sum of all terms in \eqref{eq:sym_prod_sum}
-that are associated with the diagram $g$.
-
-To find an explicit form of $\Q_g\p{\m w,\m X}$, we first consider all
-distinct ways to assign $\abs{g}$ indices to specific tensor factors
-of $\m w$ and $\m X$ in a manner consistent with $g$.  For example, if
-$\m X=\p{X_1,X_2}$ with dimensions $\p{M_1,M_2}=\p{2,1}$ and
-$\abs{g}=2$ with $g_{\set{1}}=g_{\set{1,2}}=1$, then
-$X_1\p{a,b} X_2\p{b}$ and $X_1\p{b,a} X_2\p{b}$ are distinct
-assignments of $\abs{g}$ indices to $\m X$ in a manner consistent with
-$g$, provided that $a\ne b$.  After assigning $\abs{g}$ indices to
-$\m w$ and $\m X$ in a manner consistent with $g$, we need to sum over
-all values that these indices can take, which amounts to a sum over
-all $k\in\D_N\p{\abs{g}}$.
-
-A specific assignment of indices can be thought of as a choice, for
-each of the $M_\alpha$ tensor factors of $w_\alpha$ and $X_\alpha$, of
-which dot in the diagram $g$ that tensor factor is associated with
-(i.e.~out of the $M_\alpha$ dots within circle $\alpha$).  There are
-nominally $M_\alpha!$ such choices for a given $\alpha$, and therefore
-$\prod_{\alpha\in\ZZ_p}M_\alpha!$ choices total.  The only catch is
-that these choices treat all dots in $g$ as distinct, whereas the
-$g_S$ dots within region $S$ of $g$ are actually indistinguishable.
-The assignment $X_1\p{a,b,c} X_2\p{a,b}$, for example, is the same as
-$X_1\p{b,a,c} X_2\p{b,a}$, because $a$ and $b$ are dummy indices.
-Treating $g_S$ dummy indices as as distinguishable results in
-over-counting distinct index assignments by a factor of $g_S!$.
-Denoting the set of all distinct index assignments according to $g$ by
-$\L\p{g}$, the number of such assignments is therefore
+where $\EQPS$ denotes equality under a restriction to $\M_0$, and $\Q_g\p{\m w,\m X}$ is the sum of all terms in \eqref{eq:sym_prod_sum} that are associated with the diagram $g$.
+
+To find an explicit form of $\Q_g\p{\m w,\m X}$, we first consider all distinct ways to assign $\abs{g}$ indices to specific tensor factors of $\m w$ and $\m X$ in a manner consistent with $g$.
+For example, if $\m X=\p{X_1,X_2}$ with dimensions $\p{M_1,M_2}=\p{2,1}$ and $\abs{g}=2$ with $g_{\set{1}}=g_{\set{1,2}}=1$, then $X_1\p{a,b} X_2\p{b}$ and $X_1\p{b,a} X_2\p{b}$ are distinct assignments of $\abs{g}$ indices to $\m X$ in a manner consistent with $g$, provided that $a\ne b$.
+After assigning $\abs{g}$ indices to $\m w$ and $\m X$ in a manner consistent with $g$, we need to sum over all values that these indices can take, which amounts to a sum over all $k\in\D_N\p{\abs{g}}$.
+
+A specific assignment of indices can be thought of as a choice, for each of the $M_\alpha$ tensor factors of $w_\alpha$ and $X_\alpha$, of which dot in the diagram $g$ that tensor factor is associated with (i.e.~out of the $M_\alpha$ dots within circle $\alpha$).
+There are nominally $M_\alpha!$ such choices for a given $\alpha$, and therefore $\prod_{\alpha\in\ZZ_p}M_\alpha!$ choices total.
+The only catch is that these choices treat all dots in $g$ as distinct, whereas the $g_S$ dots within region $S$ of $g$ are actually indistinguishable.
+The assignment $X_1\p{a,b,c} X_2\p{a,b}$, for example, is the same as $X_1\p{b,a,c} X_2\p{b,a}$, because $a$ and $b$ are dummy indices.
+Treating $g_S$ dummy indices as as distinguishable results in over-counting distinct index assignments by a factor of $g_S!$.
+Denoting the set of all distinct index assignments according to $g$ by $\L\p{g}$, the number of such assignments is therefore
 \begin{align}
@@
 \end{align}
-where $\m w\p{k,\ell}$ and $\m X\p{k,\ell}$ denote the scalar and
-operator acquired by assigning the indices
-$k=\p{k_1,k_2,\cdots,k_{\abs{g}}}$ to $\m w$ and $\m X$ according to
-$\ell$.  Upon projection to the PS manifold $\M_0$, this operator
-becomes
+where $\m w\p{k,\ell}$ and $\m X\p{k,\ell}$ denote the scalar and operator acquired by assigning the indices $k=\p{k_1,k_2,\cdots,k_{\abs{g}}}$ to $\m w$ and $\m X$ according to $\ell$.
+Upon projection to the PS manifold $\M_0$, this operator becomes
 \begin{align}
@@
 \end{align}
-and permutational symmetry implies that the averaging for
-$\mean{\m X}\p{\ell}$ is not actually necessary:
+and permutational symmetry implies that the averaging for $\mean{\m X}\p{\ell}$ is not actually necessary:
 \begin{align}
@@
 \end{align}
-If, furthermore, all tensors $w_\alpha$ in $\m w$ (or all operators
-$X_\alpha$ in $\m X$) are permutationally symmetric, meaning that they
-are invariant under arbitrary permutations of their tensor factors,
-then
+If, furthermore, all tensors $w_\alpha$ in $\m w$ (or all operators $X_\alpha$ in $\m X$) are permutationally symmetric, meaning that they are invariant under arbitrary permutations of their tensor factors, then
 \begin{align}
@@
 \end{align}
-In summary, the product of operators $\Q\p{\m w,\m X}$ in
-\eqref{eq:sym_prod_start} admits the diagrammatic expansion in
-\eqref{eq:sym_prod_diagrams}, as a sum of $\Q_g\p{\m w,\m X}$ over all
-diagrams $g\in\G\p{\m M}$.  Each term $\Q_g\p{\m w,\m X}$ is a sum of
-$\col{\m w}\p{\ell}\mean{\m X}\p{\ell}$ over all distinct index
-assignments $\ell\in\L\p{g}$.  The scalar $\col{\m w}\p{\ell}$ is
-obtained by assigning indices $k=\p{k_1,k_2,\cdots,k_{\abs{g}}}$ to
-the tensors $\m w$ as prescribed by $\ell$, and summing over all
-values of $k\in\D_N\p{\abs{g}}$.  The collective operator
-$\mean{\m X}\p{\ell}$ is similarly constructed with an average over
-all $k\in\D_N\p{\abs{g}}$, rather than a sum.  If the tensors $\m w$
-or operators $\m X$ are all permutationally symmetric, then
-$\Q_g\p{\m w,\m X}$ factorizes into $\col{\m w}\p{g}\mean{\m X}\p{g}$,
-where $\col{\m w}\p{g}$ and $\mean{\m X}\p{g}$ are respectively the
-sum and average of $\col{\m w}\p{\ell}$ and $\mean{\m X}\p{\ell}$ over
-all index assignments $\ell\in\L\p{g}$.
-
-As a final point, we note that the choice to sum (average) for the
-scalar (operator) content of $\Q_g\p{\m w,\m X}$ in
-\eqref{eq:sum_assignments} and \eqref{eq:sum_diagrams} was arbitrary,
-and we could just as well have chosen to interchange sums and
-averages.  To simplify the diagrammatic formalism we develop in the
-following section, we therefore define the {\it mixed} sums and
-averages
+In summary, the product of operators $\Q\p{\m w,\m X}$ in \eqref{eq:sym_prod_start} admits the diagrammatic expansion in \eqref{eq:sym_prod_diagrams}, as a sum of $\Q_g\p{\m w,\m X}$ over all diagrams $g\in\G\p{\m M}$.
+Each term $\Q_g\p{\m w,\m X}$ is a sum of $\col{\m w}\p{\ell}\mean{\m X}\p{\ell}$ over all distinct index assignments $\ell\in\L\p{g}$.
+The scalar $\col{\m w}\p{\ell}$ is obtained by assigning indices $k=\p{k_1,k_2,\cdots,k_{\abs{g}}}$ to the tensors $\m w$ as prescribed by $\ell$, and summing over all values of $k\in\D_N\p{\abs{g}}$.
+The collective operator $\mean{\m X}\p{\ell}$ is similarly constructed with an average over all $k\in\D_N\p{\abs{g}}$, rather than a sum.
+If the tensors $\m w$ or operators $\m X$ are all permutationally symmetric, then $\Q_g\p{\m w,\m X}$ factorizes into $\col{\m w}\p{g}\mean{\m X}\p{g}$, where $\col{\m w}\p{g}$ and $\mean{\m X}\p{g}$ are respectively the sum and average of $\col{\m w}\p{\ell}$ and $\mean{\m X}\p{\ell}$ over all index assignments $\ell\in\L\p{g}$.
+
+As a final point, we note that the choice to sum (average) for the scalar (operator) content of $\Q_g\p{\m w,\m X}$ in \eqref{eq:sum_assignments} and \eqref{eq:sum_diagrams} was arbitrary, and we could just as well have chosen to interchange sums and averages.
+To simplify the diagrammatic formalism we develop in the following section, we therefore define the {\it mixed} sums and averages
 \begin{align}
@@
 \end{align}
-where $\ut{\m w}\p{g}$ sums over index values $k\in\D_N\p{\abs{g}}$,
-but averages over index assignments $\ell\in\L\p{g}$; and vice versa
-for $\ot{\m X}\p{g}$.  If all tensors $w_\alpha$ in $\m w$ (or all
-operators $X_\alpha$ in $\m X$) are permutationally symmetric, then
+where $\ut{\m w}\p{g}$ sums over index values $k\in\D_N\p{\abs{g}}$, but averages over index assignments $\ell\in\L\p{g}$; and vice versa for $\ot{\m X}\p{g}$.
+If all tensors $w_\alpha$ in $\m w$ (or all operators $X_\alpha$ in $\m X$) are permutationally symmetric, then
 \begin{align}
@@
 
-Diagrams such as \eqref{eq:venn_diagram} play a central role in
-simplifying products of multi-body operators, $\Q\p{\m w,\m X}$
-through the expansion of these products in
-\eqref{eq:sym_prod_diagrams} as sum over $\Q_g\p{\m w,\m X}$.  Here,
-we formalize a diagrammatic representation of the scalars
-$\col{\m w}\p{\ell}$ and $\ut{\m w}\p{g}$ that appear in
-$\Q_g\p{\m w,\m X}$, which enables us to perform otherwise intractable
-calculations.  We then derive rules to ``reduce'' these scalars so as
-to minimize the computational cost of their numerical evaluation.
-
-A diagram $g\in\G\p{\m M}$ with $\m M=\p{M_1,M_2,\cdots,M_p}$ consists
-of $p$ circles, and $g_S$ dots at the intersection of circles
-$\alpha\in S\subset\ZZ_p$.  By slight abuse of notation, labeling each
-circle $\alpha$ by a tensor $w_\alpha$ defines the scalar
-$g\p{\m w}\equiv\ut{\m w}\p{g}$.  If, for example, all tensors
-$w_\alpha$ are all permutationally symmetric, then
+Diagrams such as \eqref{eq:venn_diagram} play a central role in simplifying products of multi-body operators, $\Q\p{\m w,\m X}$ through the expansion of these products in \eqref{eq:sym_prod_diagrams} as sum over $\Q_g\p{\m w,\m X}$.
+Here, we formalize a diagrammatic representation of the scalars $\col{\m w}\p{\ell}$ and $\ut{\m w}\p{g}$ that appear in $\Q_g\p{\m w,\m X}$, which enables us to perform otherwise intractable calculations.
+We then derive rules to ``reduce'' these scalars so as to minimize the computational cost of their numerical evaluation.
+
+A diagram $g\in\G\p{\m M}$ with $\m M=\p{M_1,M_2,\cdots,M_p}$ consists of $p$ circles, and $g_S$ dots at the intersection of circles $\alpha\in S\subset\ZZ_p$.
+By slight abuse of notation, labeling each circle $\alpha$ by a tensor $w_\alpha$ defines the scalar $g\p{\m w}\equiv\ut{\m w}\p{g}$.
+If, for example, all tensors $w_\alpha$ are all permutationally symmetric, then
 \begin{align}
@@
 \end{align}
-When the tensors $\m w$ are not all permutationally symmetric, one can
-additionally equip a diagram $g$ with a specific index assignment
-$\ell\in\L\p{g}$, in which case the diagram is equal to
-$\ell\p{\m w}\equiv\col{\m w}\p{\ell}$.  Diagrams that are not
-equipped with a specific index assignment $\ell$ are then an average
-over all $\ell\in\L\p{g}$,
-i.e.~$g\p{\m w}\equiv\sum_{\ell\in\L\p{g}} \ell\p{\m w} /
-\abs{\L\p{g}}$.  To keep mathematical expressions simple, we will use
-symmetric tensors in all examples throughout this section, but our
-discussions apply just as well to the asymmetric case.
-
-A diagram such as \eqref{eq:example_123} with $\abs{g}$ dots nominally
-takes $O\p{N^{\abs{g}}}$ time to compute due to the interdependence of
-all indices.  We wish to ``reduce'' such diagrams and minimize their
-computational cost as much as possible.  Our general strategy will be
-to replace constrained sums that make all indices interdependent by
-unconstrained sums that can be carried out independently, in sequence.
-In order to represent different constraints, we need to define
-diagrams in which filled dots (``$\bullet$'') may be replaced by empty
-dots (``$\circ$'') or crosses (``$\bm\times$'').
-
-Filled dots in a diagram correspond to indices that are constrained to
-have different values.  The presence of $F$ filled dots thus
-corresponds to a sum over $\D_N\p{F}$.  Every empty dot, meanwhile,
-corresponds to an index that is summed over $\ZZ_N$ without additional
-constraints, e.g.
+When the tensors $\m w$ are not all permutationally symmetric, one can additionally equip a diagram $g$ with a specific index assignment $\ell\in\L\p{g}$, in which case the diagram is equal to $\ell\p{\m w}\equiv\col{\m w}\p{\ell}$.
+Diagrams that are not equipped with a specific index assignment $\ell$ are then an average over all $\ell\in\L\p{g}$, i.e.~$g\p{\m w}\equiv\sum_{\ell\in\L\p{g}} \ell\p{\m w} / \abs{\L\p{g}}$.
+To keep mathematical expressions simple, we will use symmetric tensors in all examples throughout this section, but our discussions apply just as well to the asymmetric case.
+
+A diagram such as \eqref{eq:example_123} with $\abs{g}$ dots nominally takes $O\p{N^{\abs{g}}}$ time to compute due to the interdependence of all indices.
+We wish to ``reduce'' such diagrams and minimize their computational cost as much as possible.
+Our general strategy will be to replace constrained sums that make all indices interdependent by unconstrained sums that can be carried out independently, in sequence.
+In order to represent different constraints, we need to define diagrams in which filled dots (``$\bullet$'') may be replaced by empty dots (``$\circ$'') or crosses (``$\bm\times$'').
+
+Filled dots in a diagram correspond to indices that are constrained to have different values.
+The presence of $F$ filled dots thus corresponds to a sum over $\D_N\p{F}$.
+Every empty dot, meanwhile, corresponds to an index that is summed over $\ZZ_N$ without additional constraints, e.g.
 \begin{align}
@@
 \end{align}
-Crosses, meanwhile, correspond to indices that are {\it only} summed
-over the values of ``filled dot indices'', e.g.
+Crosses, meanwhile, correspond to indices that are {\it only} summed over the values of ``filled dot indices'', e.g.
 \begin{align}
@@
 \end{align}
-Empty dots and crosses allow us to decompose diagrams with a high
-computational cost into different diagrams with a lower computational
-cost.  We will essentially use two tricks to reduce diagrams:
-eliminating filled dots (which will add empty dots and crosses), and
-eliminating crosses (which will move around filled dots).  These two
-tricks will be applied in sequence until only empty dots remain.
-
-Given a diagram containing only dots (filled or empty), we can
-decompose any filled dot into an empty dot and a cross.  For example,
+Empty dots and crosses allow us to decompose diagrams with a high computational cost into different diagrams with a lower computational cost.
+We will essentially use two tricks to reduce diagrams: eliminating filled dots (which will add empty dots and crosses), and eliminating crosses (which will move around filled dots).
+These two tricks will be applied in sequence until only empty dots remain.
+
+Given a diagram containing only dots (filled or empty), we can decompose any filled dot into an empty dot and a cross.
+For example,
 \begin{align}
@@
 \end{align}
-The decomposition in \eqref{eq:example_elim} essentially breaks up a
-sum over $a\in\ZZ_N\setminus\set{b,c}$ into a difference of sums over
-$a\in\ZZ_N$ and $a\in\set{b,c}$.  Together, the sums over $a\in\ZZ_N$
-to compute all $v\p{\circ,b}$ have an $O\p{N^2}$ cost, and can be
-performed prior to evaluating the diagram in
-\eqref{eq:example_elim_o}.  The decomposition in
-\eqref{eq:example_elim} therefore splits one $O\p{N^3}$ diagram into
-two $O\p{N^2}$ diagrams.
-
-After decomposing a filled dot into an empty dot and a cross, the next
-step is to immediately eliminate the cross.  To eliminate a cross from
-a diagram, we first note that the corresponding index can ignore other
-indices on tensors that contain the index, which is to say that we can
-replace the sum over $a\in\set{b,c}$ in \eqref{eq:example_elim_x} by a
-sum over $a\in\set{c}$.  For each remaining index addressed by a
-cross, meanwhile, the cross simply forces the corresponding indices to
-be equal, which is equivalent to moving a filled dot into a different
-region:
+The decomposition in \eqref{eq:example_elim} essentially breaks up a sum over $a\in\ZZ_N\setminus\set{b,c}$ into a difference of sums over $a\in\ZZ_N$ and $a\in\set{b,c}$.
+Together, the sums over $a\in\ZZ_N$ to compute all $v\p{\circ,b}$ have an $O\p{N^2}$ cost, and can be performed prior to evaluating the diagram in \eqref{eq:example_elim_o}.
+The decomposition in \eqref{eq:example_elim} therefore splits one $O\p{N^3}$ diagram into two $O\p{N^2}$ diagrams.
+
+After decomposing a filled dot into an empty dot and a cross, the next step is to immediately eliminate the cross.
+To eliminate a cross from a diagram, we first note that the corresponding index can ignore other indices on tensors that contain the index, which is to say that we can replace the sum over $a\in\set{b,c}$ in \eqref{eq:example_elim_x} by a sum over $a\in\set{c}$.
+For each remaining index addressed by a cross, meanwhile, the cross simply forces the corresponding indices to be equal, which is equivalent to moving a filled dot into a different region:
 \begin{align}
@@
 \end{align}
-If there are multiple filled dots for the cross to eliminate, then the
-result is a sum over all choices of filled dots:
+If there are multiple filled dots for the cross to eliminate, then the result is a sum over all choices of filled dots:
 \begin{align}
@@
 \end{align}
-We can generally repeat the above process of sequentially eliminating
-filled dots and crosses until only empty dots remain, at which point
-we have made all indices in a diagram independent.
+We can generally repeat the above process of sequentially eliminating filled dots and crosses until only empty dots remain, at which point we have made all indices in a diagram independent.
 
@@
 
-The process of constructing and simplifying diagrams to compute scalar
-coefficients of the expansions in \eqref{eq:sym_prod_diagrams} and
-\eqref{eq:sum_diagrams_alt} is systematic enough to perform
-algorithmically.  The last step necessary to automate the evaluation
-of these expansions is the projection of multi-spin operators
-$\ot{\m X}\p{\ell}$ onto the PS manifold $\M_0$.  Without loss of
-generality, we consider a single $M$-spin operator $X$.  Our task is
-essentially to find coefficients of the expansion
+The process of constructing and simplifying diagrams to compute scalar coefficients of the expansions in \eqref{eq:sym_prod_diagrams} and \eqref{eq:sum_diagrams_alt} is systematic enough to perform algorithmically.
+The last step necessary to automate the evaluation of these expansions is the projection of multi-spin operators $\ot{\m X}\p{\ell}$ onto the PS manifold $\M_0$.
+Without loss of generality, we consider a single $M$-spin operator $X$.
+Our task is essentially to find coefficients of the expansion
 \begin{align}
@@
 \end{align}
-where $\A_n\p{N}$ is the set of all ways to assign $N$ (identical)
-spins to $n$ (distinct) states, such that for any $a\in\A_n\p{N}$, the
-state $\ket{a}=\ket{a_1,a_2,\cdots,a_n}$ is labeled by the occupation
-number $a_\mu$ of state $\mu$, with $\sum_\mu a_\mu = N$.  Written out
-explicitly,
+where $\A_n\p{N}$ is the set of all ways to assign $N$ (identical) spins to $n$ (distinct) states, such that for any $a\in\A_n\p{N}$, the state $\ket{a}=\ket{a_1,a_2,\cdots,a_n}$ is labeled by the occupation number $a_\mu$ of state $\mu$, with $\sum_\mu a_\mu = N$.
+Written out explicitly,
 \begin{align}
@@
 \end{align}
-where the multinomial coefficient $\C\p{a}$ counts the number of
-distinct ways to permute the tensor factors of $\ket{\tilde a}$, and
-enforces $\bk{a|a}=1$.  Using these states, we can expand
+where the multinomial coefficient $\C\p{a}$ counts the number of distinct ways to permute the tensor factors of $\ket{\tilde a}$, and enforces $\bk{a|a}=1$.
+Using these states, we can expand
 \begin{align}
@@
 \end{align}
-where the restriction $\alpha\le a$ and difference $a-\alpha$ are
-evaluated element-wise, i.e.~$\alpha\le a\implies \alpha_\mu\le a_\mu$
-and $\p{a-\alpha}_\mu=a_\mu-\alpha_\mu$ for all $\mu$; and
-$\delta_{cd}=1$ if $c=d$ and zero otherwise.  We sum over both
-$\alpha$ and $\beta$ above merely to keep the expression symmetric
-with respect to transposition; in practice, one can simply sum over
-$\alpha\in\A_n\p{M}$ and set $\beta=b-a+\alpha$ (throwing out terms
-with any $\beta_\mu<0$).  Note that, by slight abuse of notation, the
-operator $X$ on the left of \eqref{eq:multi_body_eval} acts on an
-arbitrary choice of $M$ spins (out of $N$), whereas the operator $X$
-on the right of \eqref{eq:multi_body_eval} is simply an $M$-spin
-operator, with matrix elements $\bk{\alpha|X|\beta}$ evaluated with
-respect to the $M$-spin states $\ket\alpha,\ket\beta$.
+where the restriction $\alpha\le a$ and difference $a-\alpha$ are evaluated element-wise, i.e.~$\alpha\le a\implies \alpha_\mu\le a_\mu$ and $\p{a-\alpha}_\mu=a_\mu-\alpha_\mu$ for all $\mu$; and $\delta_{cd}=1$ if $c=d$ and zero otherwise.
+We sum over both $\alpha$ and $\beta$ above merely to keep the expression symmetric with respect to transposition; in practice, one can simply sum over $\alpha\in\A_n\p{M}$ and set $\beta=b-a+\alpha$ (throwing out terms with any $\beta_\mu<0$).
+Note that, by slight abuse of notation, the operator $X$ on the left of \eqref{eq:multi_body_eval} acts on an arbitrary choice of $M$ spins (out of $N$), whereas the operator $X$ on the right of \eqref{eq:multi_body_eval} is simply an $M$-spin operator, with matrix elements $\bk{\alpha|X|\beta}$ evaluated with respect to the $M$-spin states $\ket\alpha,\ket\beta$.
 
@@
 \end{align}
-where $i,j$ index individual spins that the single-spin operators
-$X,Y$ act on; and $\m m\equiv\sum_i m_i\ket{i}$ for
-$\m m\in\set{\m v,\m w}$ are vectors of scalar coefficients.  This
-product appears, for example, in the calculation of the second-order
-effective Hamiltonian $H_1^{(2)}$ induced on the permutationally
-symmetric manifold $\M_0$ by a single-body perturbation.  Using the
-rules that we worked out in Appendix \ref{sec:diagrams} for reducing
-diagrams, we can simplify
+where $i,j$ index individual spins that the single-spin operators $X,Y$ act on; and $\m m\equiv\sum_i m_i\ket{i}$ for $\m m\in\set{\m v,\m w}$ are vectors of scalar coefficients.
+This product appears, for example, in the calculation of the second-order effective Hamiltonian $H_1^{(2)}$ induced on the permutationally symmetric manifold $\M_0$ by a single-body perturbation.
+Using the rules that we worked out in Appendix \ref{sec:diagrams} for reducing diagrams, we can simplify
 \begin{align}
@@
 \end{align}
-where $\col{m} \equiv \sum_i m_i$ for both $\m m\in\set{\m v,\m w}$,
-so
+where $\col{m} \equiv \sum_i m_i$ for both $\m m\in\set{\m v,\m w}$, so
 \begin{align}
@@
 \end{align}
-In order to write this result in terms of collective operators
-$\col{Z} \equiv \sum_i Z_i$, we expand
+In order to write this result in terms of collective operators $\col{Z} \equiv \sum_i Z_i$, we expand
 \begin{align}
@@
 \end{align}
-where $i,j,k,\ell$ index individual spins that the two-spin operators
-$X,Y$ act on; and $\m m\equiv\sum_{i,j}m_{ij}\op{i}{j}$ for
-$\m m\in\set{\m v,\m w}$ are matrices of scalar coefficients, with
-$m_{ii}=0$.  This product appears, for example, in the calculation of
-the second-order effective Hamiltonian $H_2^{(2)}$ induced on the
-permutationally symmetric manifold $\M_0$ by a two-body perturbation.
+where $i,j,k,\ell$ index individual spins that the two-spin operators $X,Y$ act on; and $\m m\equiv\sum_{i,j}m_{ij}\op{i}{j}$ for $\m m\in\set{\m v,\m w}$ are matrices of scalar coefficients, with $m_{ii}=0$.
+This product appears, for example, in the calculation of the second-order effective Hamiltonian $H_2^{(2)}$ induced on the permutationally symmetric manifold $\M_0$ by a two-body perturbation.
 The diagrammatic expansion of this operator is
@@
 \end{align}
-where $\spp{X,Y}_q$ denotes a sum over all $q$-spin operators that can
-be constructed from $X$ and $Y$, i.e.
+where $\spp{X,Y}_q$ denotes a sum over all $q$-spin operators that can be constructed from $X$ and $Y$, i.e.
 \begin{align}
@@
 \end{align}
-To minimize the computational cost of the diagrams in
-\eqref{eq:two_prod_pair}, we reduce
+To minimize the computational cost of the diagrams in \eqref{eq:two_prod_pair}, we reduce
 \begin{align}
@@
 \end{align}
-where the middle diagram in \eqref{eq:mid_zero} is zero from the fact
-that $w_{ii}=0$, and
+where the middle diagram in \eqref{eq:mid_zero} is zero from the fact that $w_{ii}=0$, and
 \begin{align}
@@
 \end{align}
-In the special case that $X=Y=Z\otimes Z$ with a single-spin operator
-$Z$ for which $Z^2=1$, we can simplify
+In the special case that $X=Y=Z\otimes Z$ with a single-spin operator $Z$ for which $Z^2=1$, we can simplify
 \begin{align}
@@
 % \end{align}
-% where $i,j,k$ index individual spins that the single-spin operators
-% $X,Y,K$ act on; and $\m m\equiv\sum_i m_i\ket{i}$ for
-% $\m m\in\set{\m u,\m v,\m w}$ are vectors of scalar coefficients.
-% This product appears, for example, in the calculation of matrix
-% elements of the operator $\O\p{\m v,Y}$ within the single-magnon
-% manifold.  The diagrammatic expansion of this operator is
+% where $i,j,k$ index individual spins that the single-spin operators $X,Y,K$ act on; and $\m m\equiv\sum_i m_i\ket{i}$ for $\m m\in\set{\m u,\m v,\m w}$ are vectors of scalar coefficients.
+% This product appears, for example, in the calculation of matrix elements of the operator $\O\p{\m v,Y}$ within the single-magnon manifold.
+% The diagrammatic expansion of this operator is
 % \begin{multline}
@@
 \end{align}
-projected onto the PS manifold $\M_0$, where $i,j,k,\ell,p,q$ index
-individual spins; $Z$ is a single-spin Ising-like operator, with
-$Z^2=1$; and $\m m\equiv\sum_{i,j} m_{ij}\op{i}{j}$ for
-$\m m\in\set{\m u,\m v,\m w}$ are matrices of scalar coefficients,
-with $m_{ii}=0$.  This product appears, for example, in the projection
-of Ising interactions onto the two-magnon manifold $\M_2$.  Collecting
-terms according to the number of $Z$ operators that remain after
-accounting for $Z^2=1$, we find that
+projected onto the PS manifold $\M_0$, where $i,j,k,\ell,p,q$ index individual spins; $Z$ is a single-spin Ising-like operator, with $Z^2=1$; and $\m m\equiv\sum_{i,j} m_{ij}\op{i}{j}$ for $\m m\in\set{\m u,\m v,\m w}$ are matrices of scalar coefficients, with $m_{ii}=0$.
+This product appears, for example, in the projection of Ising interactions onto the two-magnon manifold $\M_2$.
+Collecting terms according to the number of $Z$ operators that remain after accounting for $Z^2=1$, we find that
 \begin{align}
@@
 \end{align}
-where a diagram with unlabeled circles denotes a sum over all distinct
-label assignments, e.g.
+where a diagram with unlabeled circles denotes a sum over all distinct label assignments, e.g.
 \begin{align}
@@
 \end{align}
-The last diagram in each of $A_6,A_4,A_2,A_0$ as written in
-\eqref{eq:triple_64}, \eqref{eq:triple_2} thus has only one assignment
-of labels, as in \eqref{eq:triple_uvw}, while all other unlabeled
-diagrams have three assignments, as in \eqref{eq:triple_uvw_3}.
-
-Having identified diagrammatic representations of the coefficients
-$A_6,A_4,A_2,A_0$, we now need to simplify all diagrams to reduce
-their computational complexity, as discussed in Appendix
-\ref{sec:diagrams}.  The process of eliminating filled dots and
-crosses from a diagram is algorithmic enough to execute on a computer,
-doing which we find
+The last diagram in each of $A_6,A_4,A_2,A_0$ as written in \eqref{eq:triple_64}, \eqref{eq:triple_2} thus has only one assignment of labels, as in \eqref{eq:triple_uvw}, while all other unlabeled diagrams have three assignments, as in \eqref{eq:triple_uvw_3}.
+
+Having identified diagrammatic representations of the coefficients $A_6,A_4,A_2,A_0$, we now need to simplify all diagrams to reduce their computational complexity, as discussed in Appendix \ref{sec:diagrams}.
+The process of eliminating filled dots and crosses from a diagram is algorithmic enough to execute on a computer, doing which we find
 \begin{align}
@@
 \end{align}
-If $m_{ij}=m_{ji}$ for all $\m m\in\set{\m u,\m v,\m w}$, then we can
-define
+If $m_{ij}=m_{ji}$ for all $\m m\in\set{\m u,\m v,\m w}$, then we can define
 \begin{align}
@@
 \end{multline}
-Finally, we can also write the product $\O\p{\m u,\m v,\m w,Z}$ in
-terms of the collective operator $\col{Z} \equiv \sum_i Z_i$:
+Finally, we can also write the product $\O\p{\m u,\m v,\m w,Z}$ in terms of the collective operator $\col{Z} \equiv \sum_i Z_i$:
 \begin{align}

### 2021-04-26 sun_phases/spin_model_paper/main.tex
updates to paper

@@
 We show that in an appropriate parameter regime this system can be mapped onto a multilevel spin model with all-to-all SU($n$)-symmetric couplings.
-Additionally, Raman pulses that address internal spin states modify the atomic dispersion relation and induce spin-orbit coupling, which can act as a synthetic inhomogeneous magnetic field that competes with the SU($n$) exchange interactions.
-We investigate the mean-field dynamical phase diagram of the resulting model as a function of $n$ and different initial configurations that are accessible with the Raman pulses.
+Raman pulses that address internal spin states modify the atomic dispersion relation and induce spin-orbit coupling, which can act as a synthetic inhomogeneous magnetic field that competes with the SU($n$) exchange interactions.
+We investigate the mean-field dynamical phase diagram of the resulting model as a function of $n$ and different initial configurations that are accessible with Raman pulses.
 Consistent with previous studies for $n=2$, we find that for some initial states the spin model exhibits two distinct dynamical phases that obey simple scaling relations with $n$.
@@
 
-The tremendous theoretical significance of SU($n$) symmetries makes it all the more exciting that they appear naturally in experimental atomic, molecular, and optical (AMO) platforms \cite{wu2003exact, cazalilla2009ultracold, gorshkov2010twoorbital, beverland2016realizing, cazalilla2014ultracold, taie2012su, hofrichter2016direct, cappellini2014direct, scazza2014observation, zhang2014spectroscopic, goban2018emergence, perlin2019effective} with exquisite degrees of microscopic control.
-As a simple example, this symmetry arises through the independence of atomic orbital and interaction parameters on the $n$ nuclear spin states of alkaline-earth(-like) atoms, with e.g.~$n=10$ for ${}^{87}$Sr \cite{wu2003exact, cazalilla2009ultracold, cazalilla2014ultracold}.
+The tremendous theoretical significance of SU($n$) symmetries makes it all the more exciting that they appear naturally in experimental atomic, molecular, and optical (AMO) platforms with exquisite degrees of microscopic control.
+This symmetry arises through the independence of atomic orbital and interaction parameters on the $n$ nuclear spin states of alkaline-earth(-like) atoms, with e.g.~$n=10$ for ${}^{87}$Sr \cite{wu2003exact, cazalilla2009ultracold, gorshkov2010twoorbital, cazalilla2014ultracold}.
 As a result, AMO experiments can directly probe the role of SU($n$) interactions in controllable settings.
-Recent progress includes the experimental observations of bosonization and faster thermalization in multi-component Fermi gas with $SU(n)$ symmetry \cite{song2020evidence, sonderhouse2020thermodynamics}, SU($n$) Hubbard phases and phase transitions \cite{taie2012su, hofrichter2016direct, taie2020observation}, two-orbital SU($n$) magnetism \cite{cappellini2014direct, scazza2014observation, zhang2014spectroscopic}, and multi-body SU($n$)-symmetric interactions \cite{goban2018emergence, perlin2019effective}.
+Recent progress includes the experimental observations of bosonization and faster thermalization in multi-component Fermi gas with SU($n$) symmetry \cite{hazzard2012hightemperature, messio2012entropy, bonnes2012adiabatic, stellmer2013degenerate, yip2014theory, pagano2014onedimensional, choudhury2020collective, song2020evidence, sonderhouse2020thermodynamics}, SU($n$) Hubbard phases and phase transitions \cite{taie2012su, hofrichter2016direct, taie2020observation}, two-orbital SU($n$) magnetism \cite{cappellini2014direct, scazza2014observation, zhang2014spectroscopic, beverland2016realizing}, and multi-body SU($n$)-symmetric interactions \cite{goban2018emergence, perlin2019effective}.
+
 In the spirit of quantum simulation, further investigations in controlled settings will play an important role in understanding the consequence of SU($n$) symmetries for fundamental questions in physics, as well as their practical use in technological applications.
@@
 Working at ultracold temperatures and unit spatial filling (one atom per lattice site), we begin by mapping the SU($n$) Hubbard model onto a multilevel spin model with all-to-all SU($n$)-symmetric interactions in Section \ref{sec:spin_model}.
-We consider the effect of spin-orbit coupling (SOC) in Section \ref{sec:SOC}, finding in particular that the weak-SOC limit generally gives rise to a (synthetic) inhomogeneous magnetic field, extending previously known results to $n>2$ \cite{wall2016synthetic, livi2016synthetic, kolkowitz2016spinorbitcoupled, bromley2018dynamics, he2019engineering}.
-In Section \ref{sec:controls} we examine the use of control fields to address nuclear spins (and generate SOC), finding a simple three-laser driving scheme that allows for the preparation of interesting states with nontrivial intra-spin correlations when $n>2$.
+In Section \ref{sec:controls} we consider the use of control fields to address nuclear spins, finding a simple three-laser driving scheme that allows for the preparation of interesting states with nontrivial intra-spin correlations when $n>2$.
+We consider the effect of spin-orbit coupling (SOC) induced by control fields in Section \ref{sec:SOC}, finding in particular that the weak-SOC limit generally gives rise to a (synthetic) inhomogeneous magnetic field, extending previously known results to $n>2$ \cite{mancini2015observation, wall2016synthetic, livi2016synthetic, kolkowitz2016spinorbitcoupled, bromley2018dynamics, he2019engineering}.
 Finally, we combine these ingredients to examine mean-field dynamical behaviors of the SU($n$) spin model in Section \ref{sec:mean_field}, finding that:
@@
 
-Here we review the realization of a collective SU($n$) spin model in a system of ultracold alkaline-earth(-like) atoms trapped in an optical lattice.
+Here we derive a collective SU($n$) spin model for a system of ultracold alkaline-earth(-like) atoms trapped in an optical lattice.
 Without external driving fields, the evolution of such atoms in their electronic ground state is governed by the single-body kinetic and two-body interaction Hamiltonians
@@
   \v{\hat{A}} \cdot \v{\hat{B}}
-  \equiv \sum_{\mu,\nu} A_{\mu\nu}^\dag B_{\mu\nu}.
+  \equiv \sum_{\mu,\nu} \hat A_{\mu\nu}^\dag \hat B_{\mu\nu}.
   \label{eq:dot_product}
@@
   = -\f{u}{2N}\v\S\cdot\v\S,
-  &&
-  \v\S \equiv \sum_q \v\s_q.
   \label{eq:H_int}
 \end{align}
-Here $\v\S$ is a collective spin matrix, analogous to the collective spin vector $\vec S=(\S_\x,\S_\y,\S_\z)$ in the case of SU($2$) \cite{he2019engineering}, with $\frac12\v\S\cdot\v\S \simeq \vec S\cdot\vec S = \S_\x^2 + \S_\y^2 + \S_\z^2$ when $n=2$ (here $\simeq$ denotes equality after modding out identity terms).
+where $\v\S\equiv\sum_q\v\s_q$ is a collective spin matrix, analogous to the collective spin vector $\vec S=(\S_\x,\S_\y,\S_\z)$ in the case of SU($2$) \cite{he2019engineering}, with $\frac12\v\S\cdot\v\S \simeq \vec S\cdot\vec S = \S_\x^2 + \S_\y^2 + \S_\z^2$ when $n=2$ (here $\simeq$ denotes equality up to identity terms).
 
@@
 In this sense, $\v\s_p\cdot\v\s_q$ is analogous to the enforcement of SU($2$) spin alignment by ferromagnetic interactions, which similarly assigns different energies to the anti-symmetric spin-0 singlet $\ket{\up\dn}-\ket{\dn\up}$ and the symmetric spin-1 triplets $\set{\ket{\up\up},\ket{\dn\dn},\ket{\up\dn}+\ket{\dn\up}}$.
-By summing over all pair-wise exchange terms $\v\s_p\cdot\v\s_q$, the interaction Hamiltonian $\H_{\t{int}}$ energetically enforces a permutational symmetry among all spins, opening an energy gap $u$ between the manifold of all permutationally symmetric (PS) states and the orthogonal complement of excited (e.g.~spin-wave) states that break permutational symmetry.
+By summing over all pair-wise exchange terms $-\v\s_p\cdot\v\s_q$, the interaction Hamiltonian $\H_{\t{int}}$ energetically enforces a permutational symmetry among all spins, opening an energy gap $u$ between the manifold of all permutationally symmetric (PS) states and the orthogonal complement of excited (e.g.~spin-wave) states that break permutational symmetry.
 See Figure \ref{fig:spin_model} for a summary of this section thus far.
@@
 The value $Q_\psi\p{\v v}$ at a point $\v v\in\SS^2$ equal to the overlap of $\ket\psi$ with a pure state $\ket{\v v}$ maximally polarized along $\v v$: $Q_\psi\equiv\abs{\bk{\v v|\psi}}^2$ (see Figure \ref{fig:spin_dist}).
-In the case of a mixed state $\hat\rho$, this distribution is defined by $Q_{\hat\rho}\equiv\bk{\v v|\hat\rho|\v v}$\footnote{The dimension $n$ is reflected in $Q_{\hat\rho}$ by the fact that it can be decomposed into spherical harmonics $Y_{\ell m}$ with degree $\ell<n$ (that is, $Q_{\hat\rho}$ is ``low resolution'' or ``band-limited'' at degree $n$) \cite{perlin2020qudit}, as well as in the normalization $\int_{\v v\in\SS^2}\dd^2\v v\,Q_{\hat\rho}\p{\v v}=\frac{4\pi}{d}$.}.
+In the case of a mixed state $\hat\rho$, this distribution is defined by $Q_{\hat\rho}\equiv\bk{\v v|\hat\rho|\v v}$\footnote{The dimension $n$ is reflected in $Q_{\hat\rho}$ by the fact that it can be decomposed into spherical harmonics $Y_{\ell m}$ with degree $\ell<n$ (that is, $Q_{\hat\rho}$ is ``low resolution'' or ``band-limited'' at degree $n$) \cite{perlin2020qudit}, as well as in the normalization $\int_{\v v\in\SS^2}\dd\v v\,Q_{\hat\rho}\p{\v v}=\frac{4\pi}{d}$.}.
 Closely related spherical representations of multilevel spin states (and operators) are discussed in Ref.~\cite{li2013weylwignermoyal}.
@@
 The Hilbert space of a 10-level (spin-$\frac92$) nucleus of a ${}^{87}$Sr atom, for example, can be thought of as the Dicke manifold of 9 spin-$\frac{1}{2}$ particles (``excess'' protons and neutrons).
-We will freely use this correspondence to import standard results and constructions for the Dicke manifold (e.g.~the construction of spin operators $\s_\x,\s_\y,\s_\z$) into the context of a multilevel spin.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Spin-orbit coupling and gauge transformation}
-\label{sec:SOC}
-
-We wish to study the dynamical behavior of experimentally realizable SU($n$) spin models.
-In order to ensure validity of the $\v\S\cdot\v\S$ spin Hamiltonian in Eq.~\eqref{eq:H_int}, we consider initial preparation of a spin-polarized state\footnote{We remind the reader that validity of the spin model requires that each momentum mode is at most singly-occupied, which is guaranteed for a spin-polarized state due to Pauli blocking.}, which has the added benefit of being simple to prepare experimentally.
-However, a spin polarized state is an eigenstate of the $\v\S\cdot\v\S$ spin Hamiltonian, so we need additional ingredients for nontrivial behavior.
-Here, we consider the addition of spin-orbit coupling (SOC) induced by external driving fields.
-
-Before discussing SOC for $n$-level fermions, we first consider the well-studied case of two-level SOC with a one-dimensional lattice \cite{wall2016synthetic, kolkowitz2016spinorbitcoupled, bromley2018dynamics, he2019engineering}.
-In this case, SOC is induced by an external driving field that imprints a phase $e^{-\i\phi j}$ on lattice site $j$, or equivalently imparts a momentum kick $q\to q+\phi$, upon the absorption of a photon\footnote{In order for the drive Hamiltonian $\H_{\t{drive}}^{(\phi)}$ to be well-defined, $\phi$ should be commensurate with the lattice, with e.g.~$\phi\in2\pi/L\times\ZZ$ on a one-dimensional lattice of $L$ sites.}:
-\begin{align}
-  \H_{\t{drive}}^{(\phi)}
-  = \f{\Omega}{2} \sum_q \c_{q+\phi,\up}^\dag \c_{q,\dn} + \t{h.c.}.
-  \label{eq:drive_2}
-\end{align}
-Identifying a numerical spin index $\mu=+\frac12$ ($-\frac12$) with the state $\up$ ($\dn$), this drive Hamiltonian can be diagonalized in its momentum index $q$ by the gauge transformation $\c_{q\mu}^\dag\to \c_{q-\mu\phi,\mu}^\dag$ (equivalently $\c_{j\mu}^\dag\to e^{\i\phi\mu j} \c_{j\mu}^\dag$), which takes
-\begin{align}
-  \H_{\t{drive}}^{(\phi)} \to \H_{\t{drive}} \equiv \Omega \S_\x,
-  &&
-  \S_\x \equiv \sum_q \s_{\x,q},
-  \label{eq:drive_trans}
-\end{align}
-where $\S_\x$ is a collective spin-x operator, and $\s_{\x,q}=\frac12 \c_{q,\up}^\dag \c_{q,\dn} + \t{h.c.}$ for two-level spins.
-
-The two-level SOC drive in Eq.~\eqref{eq:drive_2} is typically implemented by coupling two electronic states of an atom with an external laser \cite{wall2016synthetic, livi2016synthetic, kolkowitz2016spinorbitcoupled, bromley2018dynamics, he2019engineering}.
-Though the coupling between optical laser fields and nuclear spins is generally too weak to drive nuclear spin transitions directly, it will nonetheless be useful to consider the $n$-level generalization of the two-level drive.
-This generalization will capture the important features of SOC for $n$-level spins, while avoiding complications that we defer to Section \ref{sec:controls}.
-
-The spin-$s$ generalization of Eq.~\eqref{eq:drive_2} is
-\begin{align}
-  \H_{\t{drive}}^{(\phi)} = \Omega_0 \sum_{q,\mu} \bk{s,\mu;1,1|s,\mu+1}
-  \c_{q+\phi,\mu+1}^\dag \c_{q\mu} + \t{h.c.},
-\end{align}
-where $\mu\in\set{s,s-1,\cdots,-s}$, and $\bk{s,\mu;1,1|s,\mu+1}$ is a Clebsch-Gordan coefficient.
-The gauge gauge transformation $\c_{q\mu}^\dag\to \c_{q-\mu\phi,\mu}^\dag$ again takes $\H_{\t{drive}}^{(\phi)}\to \Omega \S_\x$ as in Eq.~\eqref{eq:drive_trans}, where now
-\begin{align}
-  \Omega = \f{\Omega_0}{\xi},
-  &&
-  \xi \equiv -\sqrt{\f{s(s+1)}{2}},
-\end{align}
-and
-\begin{align}
-  \s_{\x,q} \equiv \xi \sum_\mu \bk{s,\mu;1,1|s,\mu+1} \c_{q,\mu+1}^\dag \c_{q\mu} + \t{h.c.}
-  \label{eq:spin_x_op}
-\end{align}
-is an $n$-level spin-x operator for spin $q$, with which $e^{-\i\theta \s_{\x,q}}$ rotates the state of spin $q$ about the x axis by an angle $\theta$ (literally rotating the sphere in Figure \ref{fig:spin_dist}).
-Despite complicating details, the change from two-level to $n$-level spins thus results in a drive Hamiltonian with an identical physical interpretation: after moving into an appropriate ``gauged frame'', this drive simply rotates all spins about the x axis at a rate $\Omega$.
-
-\begin{figure}
-\centering
-\includegraphics{soc_panels.pdf}
-\caption{
-Spin-orbit coupling for 2-level ({\bf a},{\bf c}) and 4-level ({\bf b},{\bf d}) spins.
-Colors indicate different spin projections $\mu$.
-In the ``lab frame'' ({\bf a},{\bf b}), kinetic energy is insensitive to spin, but the drive imparts a momentum kick $q\to q+\phi$ into any atom that absorbs a photon.
-Changing to the ``gauged frame'' ({\bf c},{\bf d}), essentially by shifting the momentum label $q$ for each spin state $\mu$, makes the drive diagonal in the momentum index, but comes at the cost of making kinetic energy spin-dependent.
-}
-\label{fig:soc_panels}
-\end{figure}
-
-Of course, spin-orbit coupling cannot be ``gauged away'' entirely.
-Making a gauge transformation to simplify the drive comes at the cost of making kinetic energy spin-dependent, taking
-\begin{align}
-  \H_{\t{kin}} \to \H_{\t{kin}}^{(\phi)}
-  \equiv -2J \sum_q \cos\p{q+\mu\phi} \s_{\mu\mu q},
-\end{align}
-as visualized in Figure \ref{fig:soc_panels}.
-To better interpret this Hamiltonian, we can write it in the form
-\begin{align}
-  \H_{\t{kin}}^{(\phi)}
-  = -2J \sum_q
-  \sp{\cos\p{q} \hat w_{+,q}^{(\phi)} - \sin\p{q} \hat w_{+,q}^{(\phi)}},
-\end{align}
-where
-\begin{align}
-  \hat w_{+,q}^{(\phi)} &\equiv \sum_\mu \cos\p{\mu\phi} \s_{\mu\mu q}, \\
-  \hat w_{-,q}^{(\phi)} &\equiv \sum_\mu \sin\p{\mu\phi} \s_{\mu\mu q}.
-\end{align}
-For two-level spins with $\mu=\pm\frac12$, $\hat w_{+,q}^{(\phi)}$ is proportional to the identity operator, and $\hat w_{+,q}^{(\phi)}=2\sin\p{\phi/2} \s_{\z,q}$, where $\s_{\z,q}=\sum_\mu \mu\, \s_{\mu\mu q}$ is a spin-z operator for spin $q$, so the kinetic Hamiltonian in the gauged frame describes a (synthetic) inhomogeneous magnetic field:
-\begin{align}
-  \left. \H_{\t{kin}}^{(\phi)} \right|_{n=2}
-  = 4\sin\p{\phi/2} J \sum_q \sin\p{q} \s_{\z,q}.
-\end{align}
-When $n>2$, an inhomogeneous magnetic field is likewise recovered in the limit of weak SOC, $s\phi\ll1$, in which case
-\begin{align}
-  \left. \H_{\t{kin}}^{(\phi)} \right|_{s\phi\ll1}
-  = 2J\phi \sum_q \sin\p{q} \s_{\z,q} + O\p{(s\phi)^2}.
-\end{align}
-For larger $\phi$, this Hamiltonian acquires terms with higher powers of $\s_{\z,q}$, up to $\s_{\z,q}^{n-1}$.
-
-In principle, making the gauge transformation $\c_{q\mu}^\dag\to \c_{q-\mu\phi,\mu}^\dag$ also transforms the interaction Hamiltonian as $\H_{\t{int}} \to \H_{\t{int}}^{(\phi)}$, where
-\begin{align}
-  \H_{\t{int}}^{(\phi)}
-  \equiv -\f{u}{2N} \sum_{p,q,\mu,\nu}
-  \c_{p-\mu\phi,\mu}^\dag \c_{p-\nu\phi,\nu}
-  \c_{q-\nu\phi,\nu}^\dag \c_{q-\mu\phi,\mu}.
-\end{align}
-However, $\H_{\t{int}}^{(\phi)}$ cannot be exactly reduced to a spin Hamiltonian that addresses the same spin degrees of freedom as $\H_{\t{kin}}^{(\phi)}$.
-Nonetheless, for sufficiently weak SOC (small $s\phi$) the effect of $\H_{\t{int}}^{(\phi)}$ may be approximated by $\H_{\t{int}}^{(0)}=\H_{\t{int}}$, which has previously been benchmarked for SU($2$)-symmetric interactions \cite{he2019engineering, smale2019observation}.
-To ensure that $\H_{\t{kin}}^{(\phi)}$ does not become trivial as $\phi\to0$, we can keep $J\phi/u$ constant (either by increasing $J/U$ or decreasing $N$).
-Altogether, the interacting spin Hamiltonian in the gauged frame becomes
-\begin{align}
-  \H_\spin = -\f{u}{2N} \v\S\cdot\v\S + 2J\phi \sum_q \sin\p{q} \s_{\z,q},
-  \label{eq:H_spin}
-\end{align}
-consisting of a spin-locking $\v\S\cdot\v\S$ term that energetically favors permutational symmetry, and an inhomogeneous magnetic field that causes inter-spin dephasing.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Control fields, observables, and initial states}
+We will freely use this correspondence to import standard results and constructions for the Dicke manifold (e.g.~the spin operators $\s_\x,\s_\y,\s_\z$) into the context of a multilevel spin.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{External control fields}
 \label{sec:controls}
 
-We now take a closer look at external control fields, which determine the observables we can access and the initial states we can prepare.
-To this end, we consider addressing the atoms with several lasers detuned by $\Delta$ below an electronic ${^1}\t{S}_0\to {^3}\t{P}_0$ transition of individual alkaline-earth-like atoms on a one-dimensional lattice with tight transverse confinement.
+We now consider the addition of external control fields to address atoms' internal spin states, which will determine the observables we can access and initial states we can prepare.
+Specifically, we consider off-resonantly addressing an electronic $\ket\g\to\ket\e$ transition of the atoms, and then perturbatively eliminating electronic $\ket\e$ excitations to arrive at an effective ground-state Hamiltonian addressing nuclear spins.
+For simplicity, we will assume that the total spin $s$ of the ground- and excited-state (hyperfine) manifolds are the same, as e.g.~with the ${^1}\t{S}_0\to{^3}\t{P}_0$ transition of alkaline-earth-like atoms (AEAs).
+However, the results of this section (namely the general form of effective nuclear spin Hamiltonians, as well as the corresponding set of accessible observables and initial states) are the same for transitions that take $s\to s\pm1$, so in practice one is free to address the hyperfine manifolds of the ${^1}\t{S}_0\to{^3}\t{P}_1$ transition of AEAs.
+
+We assume that all driving lasers are detuned by $\Delta$ below an electronic transition.
 Each laser is indexed by an axis of propagation, $\v v$, and a drive amplitude $\Omega_{\v v,+}$ ($\Omega_{\v v,-}$) for right-circularly (left-circularly) polarized light.
-An electronic ${^1}\t{S}_0\to {^3}\t{P}_0$ transition is thus accompanied by a nuclear spin transition $\mu_{\v v}\to\mu_{\v v}\pm1$, where $\mu_{\v v}$ is the projection of nuclear spin onto the axis $\v v$.
+An electronic $\ket\g\to\ket\e$ excitation is thus accompanied by a nuclear spin transition $\mu_{\v v}\to\mu_{\v v}\pm1$, where $\mu_{\v v}$ is the projection of nuclear spin onto the axis $\v v$.
 Furthermore, a laser pointing along $\v v$ will generally imprint a phase $e^{-\i\kappa\v v\cdot\v\ell j}$ on lattice site $j$, where $\kappa$ is the wavenumber of the laser (in units with lattice spacing $a=1$) and $\v\ell$ is a fixed unit vector parallel to the lattice.
-Altogether, the multi-laser drive Hamiltonian can be written in the form
+On a one-dimensional lattice with transverse confinement, the multi-laser drive Hamiltonian is then
 \begin{align}
@@
   + \Delta \hat N_\e,
-  \label{eq:drive_full}
-\end{align}
-where $\sigma\in\set{+1,-1}$ indexes a polarization of light; $\s_{\v v\sigma j}$ is a $n$-level spin-raising ($\sigma=+1$) or spin-lowering ($\sigma=-1$) operator along axis $\v v$ for atom $j$ (clarified below); $\ket\g_j$ and $\ket\e_j$ respectively denote the ground (${^1}\t{S}_0$) and excited (${^3}\t{P}_0$) electronic states of an atom on site $j$; and $\hat N_\e=\1\otimes\sum_j\op\e_j$ is the number of excited atoms (with $\1$ the identity operator on all spin degrees of freedom).
-
-To define the spin-raising/lowering operators $\s_{\v v\sigma j}$, we first define $\s_{\x,j}$ similarly to $\s_{\x,q}$ in Eq.~\eqref{eq:spin_x_op} with $q\leftrightarrow j$.
-The spin-raising operator $\s_{+,j}$ is then twice the spin-raising part of $\s_{\x,j}$, just as in the case of SU($2$) where $\s_\x=\frac12\op{\up}{\dn}+\t{h.c.}$, and the spin-lowering operator is $\s_{-,j}\equiv \s_{+,j}^\dag$.
-The operator $\s_{\v v,+\,j}$ that raises spin along $\v v$ is acquired by a rotation of $\s_{+,j}$ that maps the north pole to $\v v$:
-\begin{align}
-  \s_{\v v,+} \equiv \hat U_{\v v} \s_+ \hat U_{\v v}^\dag,
-  &&
-  \hat U_{\v v} \equiv e^{-\i\alpha_{\v v}\s_\z} e^{-\i\beta_{\v v}\s_\y},
-  \label{eq:spin_rot}
-\end{align}
-where we have dropped the site index $j$ for brevity; $\alpha_{\v v}$ and $\beta_{\v v}$ are the azimuthal and polar angles of $\v v$; and $\s_\y\equiv-\frac\i2\s_++\t{h.c.}$.
-Note that the operators $(\s_\x,\s_\y,\s_\z)$ form an $\su(2)$ algebra, so given a particular axis $\v v$ one can expand $\s_{\v v,+}=\sum_m f_{\v vm} \s_m$ with coefficients $f_{\v vm}$ that are independent of the dimension $n$.
-In particular, one is always free to set $n=2$ and expand Eq.~\eqref{eq:spin_rot} with spin-$\frac12$ operators (or Pauli matrices) to find the coefficients $f_{\v vm}$.
+  \label{eq:H_drive_full}
+\end{align}
+where $\sigma\in\set{+1,-1}$ indexes a polarization of light; $\s_{\v v\sigma j}$ is an $n$-level spin-raising ($\sigma=+1$) or spin-lowering ($\sigma=-1$) operator along axis $\v v$ for an atom on lattice site $j$ (identical to that for $n-1$ spin-$\frac12$ particles in the Dicke manifold); $\ket\g_j$ and $\ket\e_j$ respectively denote the ground and excited electronic states of atom $j$; and $\hat N_\e=\1\otimes\sum_j\op\e_j$ counts the number of excited atoms (with $\1$ the identity operator on all spin degrees of freedom).
 
@@
 A third, linearly polarized laser with amplitude $\Omega_0$ points in a direction orthogonal to both the lattice and the other driving lasers.
-Absorbing a photon from the laser with amplitude $\Omega_m$ induces a transition $(\g,\mu)\to(\e,\mu+m)$ for the (electronic, nuclear spin) state of an atom, where nuclear spin is quantized along the axis of propagation for the $m=+1$ laser.
+Absorbing a photon from the laser with amplitude $\Omega_m$ induces a transition $(\g,\mu)\to(\e,\mu+m)$ for the (electronic, nuclear spin) state of an atom, where nuclear spin is quantized along the z axis.
 }
@@
 In the far-detuned limit $\abs{\Delta}\gg\abs{\Omega_m}$, a second-order perturbative treatment of electronic excitations ($\ket\e$) yields an effective drive Hamiltonian that only addresses ground-state nuclear spins.
-After additionally making the gauge transformation $\c_{j\mu}^\dag\to e^{\i\phi\mu j} \c_{j\mu}^\dag$ (equivalently $\s_{mj}\to e^{\i m\phi j}\s_{mj}$), the drive Hamiltonian then becomes
+After additionally making the gauge transformation $\s_{mj}\to e^{\i m\phi j}\s_{mj}$ (equivalently $\c_{j\mu}^\dag\to e^{\i\phi\mu j} \c_{j\mu}^\dag$), the drive Hamiltonian then becomes
 \begin{align}
   \H_{\t{3LD}} = \sum_j \H_{\t{3LD},j}^{\t{single}},
-  \label{eq:drive_all}
+  \label{eq:H_3LD}
 \end{align}
@@
   + \tilde\Omega_0 \tilde\Omega_- \s_\x
-  + \tilde\Omega_0 \tilde\Omega_+ \p{\s_\z \s_\x  + \s_\x \s_\z} \\
+  + \tilde\Omega_0 \tilde\Omega_+ (\s_\z \s_\x  + \s_\x \s_\z) \\
   - \tilde\Omega_0^2 \s_\z^2 - \tilde\Omega_+^2 \s_\x^2
   - \tilde\Omega_-^2 \s_\y^2,
-  \label{eq:drive_single}
+  \label{eq:H_3LD_single}
 \end{multline}
@@
 \end{align}
-where we have made the simplifying assumption that all drive amplitudes are real to arrive at the form of $\H_{\t{3LD}}^{\t{single}}$ in Eq.~\eqref{eq:drive_single}.
+where we have made the simplifying assumption that all drive amplitudes are real to arrive at the form of $\H_{\t{3LD}}^{\t{single}}$ in Eq.~\eqref{eq:H_3LD_single}.
 We relax the assumption of real drive amplitudes in Appendix \ref{sec:full_drive}.
@@
 Drive Hamiltonians (left column) that can be implemented with different amplitude-matching conditions (right column), some of which are specified by an arbitrary sign $\sigma\in\set{+1,-1}$.
-The drives shown here are equal to that Eq.~\eqref{eq:drive_single} up to a possible energy shift of $\s_\x^2+\s_\y^2+\s_\z^2=s(s+1)$, and come in mutually commuting pairs: a drive with $\Omega_m=1$ and $\Omega_n=0$ for both $n\ne m$ commutes with the drive in which $\Omega_m=0$ and both $\abs*{\Omega_n}=1$.
+The drives shown here are equal to that Eq.~\eqref{eq:H_3LD_single} up to a possible energy shift of $\s_\x^2+\s_\y^2+\s_\z^2=s(s+1)$, and come in mutually commuting pairs: a drive with $\Omega_m=1$ and $\Omega_n=0$ for both $n\ne m$ commutes with the drive in which $\Omega_m=0$ and both $\abs*{\Omega_n}=1$.
 }
@@
   \\ \hline
-  $\pm \s_\z \pm \sigma \s_\x + \sigma \p{\s_\z \s_\x + \s_\x \s_\z}$
+  $\pm \s_\z \pm \sigma \s_\x + \sigma (\s_\z \s_\x + \s_\x \s_\z)$
   & $\p{1,\sigma,\pm\sigma}$
@@
 
-Eqs.~\eqref{eq:drive_all} and \eqref{eq:drive_single} are the main technical results of this section.
-There are three important observations to make about these results.
-First, the fact that $\H_{\t{3LD}}$ acts identically on all spins means we can freely replace the site index $j$ with a momentum index $q$ (as can be verified by substituting $\c_{j\mu}=\frac1{\sqrt{L}}\sum_k e^{-\i q\cdot j} \c_{q\mu}$), which is important to ensure that this drive addresses the same spin degrees of freedom as the spin Hamiltonian $\H_\spin$ in Eq.~\eqref{eq:H_spin}.
+There are three important observations to make about Eqs.~\eqref{eq:H_3LD} and \eqref{eq:H_3LD_single}.
+First, the fact that $\H_{\t{3LD}}$ acts identically on all spins means we can freely replace the site index $j$ with a momentum index $q$ (as can be verified by substituting $\c_{j\mu}=\frac1{\sqrt{L}}\sum_k e^{-\i q\cdot j} \c_{q\mu}$), which is important to ensure that this drive addresses the same spin degrees of freedom as the spin Hamiltonians previously considered in Section \ref{sec:spin_model}.
 Second, each of $\tilde\Omega_0,\tilde\Omega_+,\tilde\Omega_-$ can be tuned independently by changing the amplitudes of the driving lasers; some particular Hamiltonians for specific values of these amplitudes are shown in Table \ref{tab:drives}.
@@
 The capability to perform arbitrary spatial rotations, together with the capability to measure the number of atoms with spin projection $\mu$ onto a fixed quantization axis, $\bk{\S_{\mu\mu}}$ (where $\S_{\mu\nu}=\sum_j\s_{\mu\nu j}$), implies the capability to reconstruct all components of the mean collective spin matrix $\bk{\v\S}=\sum_{\mu\nu}\bk{\S_{\mu\nu}}\op{\mu}{\nu}$ via spin qudit tomography \cite{newton1968measurability, perlin2020qudit}.
-Moreover, we expect that advanced quantum optimal control techniques (similar to those of Refs.~\cite{anderson2015accurate, lucarelli2018quantum}) can be used to implement arbitrary SU($n$) rotations by designing suitable time-dependent drive amplitudes $\Omega_m$.
+Moreover, we expect that advanced quantum optimal control techniques (similar to those of Refs.~\cite{anderson2015accurate, lucarelli2018quantum}) can be used to implement arbitrary SU($n$) rotations by designing suitable time-dependent drive amplitudes.
+
+If the excited-state manifold $\ket\e$ has total spin $s\pm1$, the effective ground-state Hamiltonians in Eq.~\eqref{eq:H_3LD_single} and Table \ref{tab:drives} remain almost identical, but with some additional $n$-dependent factors that do not affect the general results and discussions above.
+These results still hold if (for example) all excited hyperfine manifolds of an electronic ${^1}\t{S}_0\to{^3}\t{P}_1$ transition (with total spins $s+1,s,s-1$) are addressed simultaneously.
+See Appendix \ref{sec:full_drive} for additional details.
 
 Finally, we comment on the preparation of initial states.
-Initial states are nominally prepared in the ``lab frame'', and must be transformed according to the gauge transformation $\c_{q\mu}^\dag\to \c_{q-\mu\phi,\mu}^\dag$ prior to evolution under the ``gauged frame'' spin Hamiltonian $\H_\spin$ in Eq.~\eqref{eq:H_spin} or the three-laser drive $\H_{\t{3LD}}$ in Eq.~\eqref{eq:drive_all}.
+Initial states are nominally prepared in the ``lab frame'', and must be transformed according to the gauge transformation $\c_{j\mu}^\dag\to e^{\i\phi\mu j} \c_{j\mu}^\dag$ prior to evolution under the three-laser drive $\H_{\t{3LD}}$ in Eq.~\eqref{eq:H_3LD}, which is expressed in the ``gauge frame''.
 We assume the capability to prepare an initial state in which all spins are maximally polarized along the z axis, i.e.~$\ket{\z}^{\otimes N}=\ket{s}^{\otimes N}$, which is unaffected by the gauge transformation (up to a global phase).
-The three-laser then allows us to rotate this state into one that is polarized along any spatial axis (in the gauged frame).
+The three-laser then allows us to rotate this state into one that is polarized along any spatial axis (in the gauge frame).
 In addition, when $n>2$ the three-laser drive allows us to prepare product states with nontrivial intra-spin correlations.
-For example, when $n$ is even we can prepare an $N$-fold product of the state
+For example, when $n$ is even we can prepare an $N$-fold product of the ``kitten'' state
 \begin{align}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Spin-orbit coupling}
+\label{sec:SOC}
+
+We now consider the effect of spin-orbit coupling (SOC) induced by the control fields in Section \ref{sec:controls}.
+Before discussing SOC for $n$-level fermions, we briefly review the well-studied case of two-level SOC with a one-dimensional lattice \cite{wall2016synthetic, kolkowitz2016spinorbitcoupled, bromley2018dynamics, he2019engineering}.
+In this case, SOC is induced by an external driving field that imprints a phase $e^{-\i\phi j}$ on lattice site $j$, or equivalently imparts a momentum kick $q\to q+\phi$, upon the absorption of a photon\footnote{In order for the drive Hamiltonian $\H_{\t{drive}}^{(\phi)}$ to be well-defined, $\phi$ should be commensurate with the lattice, e.g.~$\phi\in\ZZ\times2\pi/L$ on a one-dimensional lattice of $L$ sites.}:
+\begin{align}
+  \H_{\t{drive}}^{(\phi)}
+  = \f{\Omega}{2} \sum_q \c_{q+\phi,\up}^\dag \c_{q,\dn} + \t{h.c.}.
+  \label{eq:drive_2}
+\end{align}
+Identifying a numerical spin index $\mu=+\frac12$ ($-\frac12$) with the state $\up$ ($\dn$), this drive Hamiltonian can be diagonalized in its momentum index $q$ by the gauge transformation $\c_{q\mu}^\dag\to \c_{q-\mu\phi,\mu}^\dag$ (equivalently $\c_{j\mu}^\dag\to e^{\i\phi\mu j} \c_{j\mu}^\dag$), which takes
+\begin{align}
+  \H_{\t{drive}}^{(\phi)} \to \H_{\t{drive}} \equiv \Omega \S_\x,
+  &&
+  \S_\x \equiv \sum_q \s_{\x,q},
+  \label{eq:drive_trans}
+\end{align}
+where $\s_{\x,q}=\frac12 \c_{q,\up}^\dag \c_{q,\dn} + \t{h.c.}$ for two-level spins.
+
+The two-level SOC drive in Eq.~\eqref{eq:drive_2} has been implemented with an external laser that couples the two electronic states of nuclear-spin-polarized atoms, with $\dn$ ($\up$) indexing the ground (excited) electronic state \cite{wall2016synthetic, livi2016synthetic, kolkowitz2016spinorbitcoupled, bromley2018dynamics, he2019engineering}.
+In contrast, the drive we considered in Section \ref{sec:controls} addresses electronic excitations off-resonantly, inducing an effective Hamiltonian in the ground-state hyperfine manifold with spin projections $\mu\in\set{s,s-1,\cdots,-s}$ (a similar one-laser scheme was used to study SOC using a subspace of the ground-state manifold in Ref.~\cite{mancini2015observation}).
+Nonetheless, both the two-level drive in Eq.~\eqref{eq:drive_2} and the $n$-level drive in Eq.~\eqref{eq:H_3LD} become homogeneous (i.e.~independent of the spatial mode index $j$ or $q$) and independent of the SOC angle $\phi$ after the same spin-symmetric gauge transformation\footnote{Note that the ``asymmetric'' gauge transformation $(\c_{j,\up}^\dag,\c_{j,\dn}^\dag)\to(e^{\i\phi j} \c_{j,\up}^\dag,\c_{j,\dn}^\dag)$, sometimes performed in the two-state SOC literature, does not generalize as nicely to $n>2$.} $\c_{j\mu}^\dag\to e^{\i\phi\mu j} \c_{j\mu}^\dag$.
+
+Of course, spin-orbit coupling cannot be ``gauged away'' entirely.
+Making a gauge transformation to simplify the drive comes at the cost of making the kinetic energy in Eq.~\eqref{H_kin_start} spin-dependent, taking
+\begin{align}
+  \H_{\t{kin}} \to \H_{\t{kin}}^{(\phi)}
+  \equiv -2J \sum_q \cos\p{q+\mu\phi} \s_{\mu\mu q},
+\end{align}
+as visualized in Figure \ref{fig:soc_panels}.
+To better interpret this Hamiltonian, we can write it in the form
+\begin{align}
+  \H_{\t{kin}}^{(\phi)}
+  = -2J \sum_q
+  \sp{\cos\p{q} \hat w_{+,q}^{(\phi)} - \sin\p{q} \hat w_{+,q}^{(\phi)}},
+\end{align}
+where
+\begin{align}
+  \hat w_{+,q}^{(\phi)} &\equiv \sum_\mu \cos\p{\mu\phi} \s_{\mu\mu q}, \\
+  \hat w_{-,q}^{(\phi)} &\equiv \sum_\mu \sin\p{\mu\phi} \s_{\mu\mu q}.
+\end{align}
+For two-level spins with $\mu=\pm\frac12$, $\hat w_{+,q}^{(\phi)}$ is proportional to the identity operator and $\hat w_{+,q}^{(\phi)} = 2\sin\p{\phi/2} \s_{\z,q}$, so the kinetic Hamiltonian in the gauge frame describes a (synthetic) inhomogeneous magnetic field:
+\begin{align}
+  \left. \H_{\t{kin}}^{(\phi)} \right|_{n=2}
+  = 4 J \sin\p{\phi/2} \sum_q \sin\p{q} \s_{\z,q}.
+\end{align}
+When $n>2$, an inhomogeneous magnetic field is likewise recovered in the weak SOC limit $s\phi\ll1$, in which case
+\begin{align}
+  \left. \H_{\t{kin}}^{(\phi)} \right|_{s\phi\ll1}
+  = 2J\phi \sum_q \sin\p{q} \s_{\z,q} + O\p{(s\phi)^2}.
+\end{align}
+For larger $\phi$, this Hamiltonian acquires terms with higher powers of $\s_{\z,q}$, up to $\s_{\z,q}^{n-1}$.
+
+\begin{figure}
+\centering
+\includegraphics{soc_panels.pdf}
+\caption{
+Spin-orbit coupling for 2-level ({\bf a},{\bf c}) and 4-level ({\bf b},{\bf d}) spins.
+Colors indicate different spin projections $\mu$.
+In the ``lab frame'' ({\bf a},{\bf b}), kinetic energy is insensitive to spin, but the drive imparts a momentum kick $q\to q+\phi$ into any atom that absorbs a photon.
+Changing to the ``gauge frame'' ({\bf c},{\bf d}), essentially by shifting the momentum label $q$ for each spin state $\mu$, makes the drive diagonal in the momentum index, but comes at the cost of making kinetic energy spin-dependent.
+}
+\label{fig:soc_panels}
+\end{figure}
+
+In principle, making the gauge transformation $\c_{q\mu}^\dag\to \c_{q-\mu\phi,\mu}^\dag$ also transforms the interaction Hamiltonian as $\H_{\t{int}} \to \H_{\t{int}}^{(\phi)}$, where
+\begin{align}
+  \H_{\t{int}}^{(\phi)}
+  \equiv -\f{u}{2N} \sum_{p,q,\mu,\nu}
+  \c_{p-\mu\phi,\mu}^\dag \c_{p-\nu\phi,\nu}
+  \c_{q-\nu\phi,\nu}^\dag \c_{q-\mu\phi,\mu}.
+\end{align}
+However, $\H_{\t{int}}^{(\phi)}$ cannot be exactly reduced to a spin Hamiltonian that addresses the same spin degrees of freedom as $\H_{\t{kin}}^{(\phi)}$.
+Nonetheless, for sufficiently weak SOC (small $s\phi$) the effect of $\H_{\t{int}}^{(\phi)}$ may be approximated by $\H_{\t{int}}^{(0)}=\H_{\t{int}}$, which has previously been benchmarked for SU($2$)-symmetric interactions \cite{he2019engineering, smale2019observation}.
+To ensure that $\H_{\t{kin}}^{(\phi)}$ does not become trivial as $\phi\to0$, we can keep $J\phi/u$ constant (either by increasing $J/U$ or decreasing $N$).
+Altogether, the interacting spin Hamiltonian in the gauge frame becomes
+\begin{align}
+  \H_\spin = -\f{u}{2N} \v\S\cdot\v\S + 2J\phi \sum_q \sin\p{q} \s_{\z,q},
+  \label{eq:H_spin}
+\end{align}
+consisting of a spin-locking $\v\S\cdot\v\S$ term that energetically favors permutational symmetry, and an inhomogeneous magnetic field that causes inter-spin dephasing.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Mean-field theory and dynamical phases}
@@
 
-We now return to the spin Hamiltonian $\H_\spin$ in Eq.~\eqref{eq:H_spin}, which is valid in the weak-SOC limit $s\phi\ll1$, and henceforth work exclusively in the ``gauged frame'' of $\H_\spin$ and the three-laser drive $\H_{\t{3LD}}$ in Eq.~\eqref{eq:drive_all}.
+We now study the dynamical behavior of the SOC spin Hamiltonian $H_\spin$ in Eq.~\eqref{eq:H_spin}, and henceforth work exclusively in the ``gauge frame'' of $\H_\spin$ and the three-laser drive $\H_{\t{3LD}}$ in Eq.~\eqref{eq:H_3LD}.
+The general idea will be a Ramsey-type setup in which we first prepare an initial state with the three-laser drive (using fast pule sequences), let this state evolve freely for some time under $\H_\spin$, and then again use the three-laser drive to map observables of interest onto spin projection measurements (e.g.~with spin qudit tomography \cite{newton1968measurability, perlin2020qudit}).
 At the mean-field (MF) level, up to constant energy shifts the undriven spin Hamiltonian becomes
@@
   \H_\MF = u \sum_q\sp{-\bk{\bar{\v s}}\cdot\v\s_q
-    + h \sum_q \sin\p{q} \s_{\z,q}}.
+    + h \sin\p{q} \s_{\z,q}}.
   \label{eq:H_MF}
@@
 Fixing the atom number $N$, the spin Hamiltonian has one free parameter, $h$, which determines the relative strength of the single-particle and interaction terms.
-One should therefore expect distinct dynamical behaviors when $h\ll1$, in which case strong spin-locking interactions should give rise to a long-range ordered phase, as opposed to $h\gg1$, in which case long-range order should be destroyed by the strong inhomogeneous magnetic field.
+One should therefore expect distinct dynamical behaviors when $h\ll1$, in which case strong spin-locking interactions should give rise to a long-range ordered phase, as opposed to $h\gg1$, in which case long-range order should be destroyed by the strong inhomogeneous magnetic field \cite{smale2019observation}.
 
@@
 \end{align}
-where $\bk{\O\p{t}}_\MF$ is the mean-field value of observable $\O$ at time $t$ (see Appendices \ref{sec:MFT} and \ref{sec:bosons} for details about our mean-field simulations).
+where $\bk{\O\p{t}}_\MF$ is the mean-field value of observable $\O$ at time $t$.
 Specifically, we consider the time-averaged magnetization
@@
 
+Our numerical simulations of mean-field dynamics are performed with a Schwinger boson decomposition of spin operators: $\s_{\mu\nu q} = \b_{\mu q}^\dag\b_{\nu q}$.
+This decomposition requires no approximations, and reduces the number of variables to keep track of by a factor of $\sim n$.
+See Appendices \ref{sec:MFT} and \ref{sec:bosons} additional details about our numerical simulations and the Schwinger boson equations of motion.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Spin-polarized initial state}
+\subsection{Initial spin-polarized state}
 
@@
 \caption{
-Time-averaged magnetization $\sigma_\MF$ and interaction energy $\bbk{\ss}_\MF$ for different spin dimensions $n$ (indicated in the legend) as determined by mean-field simulations of $N=100$ spins initially in the $x$-polarized state $\ket\X$ for a time $T=10^5/U$.
+Time-averaged magnetization $\sigma_\MF$ and interaction energy $\bbk{\ss}_\MF$ for different spin dimensions $n$ (indicated in the legend) as determined by mean-field simulations of $N=100$ spins initially in the x-polarized state $\ket\X$ for a time $T=10^5/u$.
 Insets show same data after rescaling $h\to h \times (n/2)^{1/3}$, and transforming vertical axes according to Eq.~\eqref{eq:rescale}.
@@
 
-Figure \ref{fig:mean_mag_int_X} shows the time-averages of the magnetization $\sigma_\MF$ and interaction energy $\bbk{\ss}_\MF$ as computed by mean-field simulations of $N=100$ spins initially in the $x$-polarized state $\ket\X\equiv\ket\x^{\otimes N}$, where
+Figure \ref{fig:mean_mag_int_X} shows the time-averages of the magnetization $\sigma_\MF$ and interaction energy $\bbk{\ss}_\MF$ as computed by mean-field simulations of $N=100$ spins initially in the x-polarized state $\ket\X\equiv\ket\x^{\otimes N}$, where
 \begin{align}
@@
 \caption{
-The critical value of $h_\crit$ as determined by mean-field simulations of $N=100$ spins initially in the $x$-polarized state $\ket\X$.
+The critical value of $h_\crit$ as determined by mean-field simulations of $N=100$ spins initially in the x-polarized state $\ket\X$.
 A single-parameter fit to $h_\crit=\p{n/2}^{-\alpha}$ finds $\alpha=0.333(5)$, and $\alpha=1/3$ is consistent with all mean-field results to within an uncertainty determined by the resolution of $h$ in mean-field simulations.
@@
 In the weak-field limit $h\to0$, the spin-locking $\v\S\cdot\v\S$ interactions of the Hamiltonian $\H_\spin$ energetically restrict dynamics to the permutationally symmetric (PS) manifold.
-At second order in perturbation theory, the effective spin Hamiltonian within the PS manifold becomes (see Appendix \ref{sec:pert_theory})
+To first order in $h$, the effect of the inhomogeneous field can be acquired by projecting it onto the PS manifold, which takes $\s_{\z,q}\to\frac1N\S_\z$.
+The first order effect of the inhomogeneous field thus vanishes, as
+\begin{align}
+  \sum_q \sin\p{q} \s_{\z,q}
+  \to \sum_q \sin\p{q} \times \frac1N \S_\z
+  = 0.
+\end{align}
+At second order in $h$, the effective Hamiltonian within the PS manifold is related to the variance of the inhomogeneous field, rather than its (vanishing) average.
+On a high level, the second-order effect of the inhomogeneous field within the PS manifold thus consists of permutation-symmetrized products of two spin-z operators, $\s_{\z,p}\s_{\z,q}$ (with $p,q$ possibly equal).
+Altogether, the effective spin Hamiltonian at second order in $h$ is (see Appendix \ref{sec:pert_theory})
 \begin{align}
@@
 The weak-field effective Hamiltonian preserves permutational symmetry, so $\bbk{\ss}_\MF\to1$ as $h\to0$.
-Moreover, both the spin Hamiltonian $\H_\spin$ and the initial state $\ket\X$ are invariant (up to global phase) under the action of $\hat\Pi_\z \R_\x^\pi$, where $\R_\x^\theta \equiv e^{-\i\theta \S_\x}$ is a rotation about the x axis by $\theta$ and $\hat\Pi_\z\equiv\sum_{q>0}\v\s_q\cdot\v\s_{-q}$ is a permutation of spins $q\leftrightarrow-q$\footnote{$\hat\Pi_\z$ can also be thought of a reflection operator that inverts momentum along the z axis.}, which is to say that
-\begin{align}
-  e^{-\i t \H_\spin} \ket\X
-  &\simeq \p{\hat\Pi_\z \R_\x^\pi e^{-\i t \H_\spin} \R_\x^\pi{}^\dag \hat\Pi_\z^\dag} \hat\Pi_\z \R_\x^\pi \ket\X
-  \label{eq:sym_x_start} \\
-  &\simeq \hat\Pi_\z \R_\x^\pi e^{-\i t \H_\spin} \ket\X,
-\end{align}
-where $\simeq$ denotes equality up to a global phase.
-The permutation $\hat\Pi_\z$ commutes with $\S_\y$, while the rotation $\R_\x^\pi$ takes $\S_\y\to-\S_\y$, so
-\begin{align}
-  \bk{\S_\y}
-  = \bk{\R_\x^\pi{}^\dag \hat\Pi_\z^\dag \S_\y \hat\Pi_\z \R_\x^\pi}
-  = -\bk{\S_\y} = 0
-  \label{eq:sym_x_end}
-\end{align}
-at all times.
-In turn, conservation of $\bk{\S_\y}=\bk{\S_\z}=0$ and permutational symmetry together imply that the magnetization $\sigma_\MF$ is determined by the time-average of $\hat\sigma_\x\equiv\s_\x/s$ for a single spin:
+Moreover, the initial y-magnetization $\bk{\S_\y}=0$ is conserved by $\H_\spin$, so long-time-averaged magnetization $\sigma_\MF$ is determined by the time-average of $\hat\sigma_\x\equiv\s_\x/s$ for a single spin:
 \begin{align}
@@
 \end{align}
-When going beyond mean-field theory, inter-spin correlations generated by $\S_\z^2$ in Eq.~\eqref{eq:H_spin_eff} will cause $\bk{\S_\x}$ to decay on time scales that diverge as $N\to\infty$, so the mean-field magnetization $\sigma_\MF$ should be interpreted as an ``intermediate-time-average'' of $\vec\sigma$.
+When going beyond mean-field theory, inter-spin correlations generated by $\S_\z^2$ in Eq.~\eqref{eq:H_spin_eff} will cause $\bk{\S_\x}$ (and thereby the magnetization $\bk{\vec\sigma}$) to decay on time scales that diverge as $N\to\infty$.
+On a lattice of linear size $L$ without periodic boundary conditions, additional corrections to the behavior predicted above will appear on $O(L/J)$ timescales.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Kitten initial states}
+\subsection{Initial kitten states}
 
@@
 A corollary of Figure \ref{fig:mean_mag_int_X} for the initial state $\ket\XX$.
-The inset for interaction energy $\bbk{\ss}_\MF$ in Figure \ref{fig:mean_mag_int_X} subtracts off the minimal value of $\bbk{\ss}_\MF$ and rescales to lie on the interval $[0,1]$, as prescribed by Eq.~\eqref{eq:rescale}.
+The inset for interaction energy $\bbk{\ss}_\MF$ in Figure \ref{fig:mean_mag_int_X} subtracts off the minimal value of $\bbk{\ss}_\MF$ and rescales to lie on the interval $[0,1]$, as prescribed in Eq.~\eqref{eq:rescale}.
 Here the subtracting and rescaling is identical, but with a minimal value of $2\gamma\p{n}$ rather than $\gamma\p{n}$.
@@
 \end{align}
-Similarly to Figure \ref{fig:mean_mag_int_X}, Figures \ref{fig:mean_mag_int_XX} and \ref{fig:mean_mag_int_XXI} show the time-averaged magnetization $\sigma_\MF$ and interaction energy $\bbk{\ss}_\MF$ throughout mean-field dynamics of the initial states $\ket\XX$ and $\ket\XXI$.
-These figures exclude the trivial case of spin dimension $n=2$, for which $\ket\xx=\ket{-s}$ is an eigenstate of $\H_\spin$ and $\ket\xxi=e^{-\i\frac{\pi}{2}\s_\z}\ket\x\equiv\ket\y$ is spin-polarized along $y$.
+Similarly to Figure \ref{fig:mean_mag_int_X}, Figures \ref{fig:mean_mag_int_XX} and \ref{fig:mean_mag_int_XXI} show the time-averaged magnetization $\sigma_\MF$ and interaction energy \mbox{$\bbk{\ss}_\MF$} throughout mean-field dynamics of the initial states $\ket\XX$ and $\ket\XXI$.
+These figures exclude the trivial case of spin dimension $n=2$, for which $\ket\xx=\ket{-s}$ is an eigenstate of $\H_\spin$ and $\ket\xxi=e^{-\i\frac{\pi}{2}\s_\z}\ket\x\equiv\ket\y$ is spin-polarized along the y axis.
 The first and perhaps most interesting observation to make about Figures \ref{fig:mean_mag_int_XX} and \ref{fig:mean_mag_int_XXI} is that they are different, signifying the importance of intra-spin coherences for the dynamical behavior of multilevel spin models.
@@
 For all initial states that we have considered, the value of $\bk{\S_\z}=0$ is conserved by the spin Hamiltonian $\H_\spin$.
-Moreover, both the spin Hamiltonian $\H_\spin$ and the state $\ket\XX$ are invariant (up to global phase) under the action of $\R_\z^\pi$, where $\R_\z^\theta\equiv e^{-\i\theta \S_\z}$.
-Similarly to Eqs.~\eqref{eq:sym_x_start}--\eqref{eq:sym_x_end}, this symmetry implies that
+Moreover, both the spin Hamiltonian $\H_\spin$ and the state $\ket\XX$ are invariant (up to global phase) under the action of $\R_\z^\pi$, where $\R_\z^\theta\equiv e^{-\i\theta \S_\z}$, which is to say that
+\begin{align}
+  \R_\z^\pi \H_\spin \R_\z^\pi{}^\dag = \H_\spin
+  &&
+  \R_\z^\pi \ket\XX \simeq \ket\XX,
+\end{align}
+where $\simeq$ denotes equality up to an overall phase.
+This symmetry implies that
 \begin{align}
@@
 Starting with an SU($n$) Hubbard model describing ultracold fermionic alkaline-earth(-like) atoms on an optical lattice, we derived a momentum-space multilevel spin model with all-to-all SU($n$)-symmetric interactions.
-We then considered the effect of spin-orbit coupling (SOC) induced by external driving lasers, finding that maintaining integrity of the spin model seems to require weak SOC, which in turn gives rise to a (synthetic) inhomogeneous magnetic field.
-Taking a closer look at the external controls that induce SOC, we found a simple three-laser drive that enables homogeneously addressing nuclear spins with a variety of spin Hamiltonians, enabling for example the preparation of ``kitten'' states in which each spin is in a coherent superposition of states spin-polarized in opposite directions.
+We then introduced external control fields, finding a simple three-laser drive that homogeneously addresses nuclear spins with a variety of spin Hamiltonians.
+Taking a closer look at the effect of the spin-orbit coupling (SOC) induced by the driving lasers, we found that maintaining integrity of the spin model requires weak SOC, which in turn gives rise to a (synthetic) inhomogeneous magnetic field.
 Finally, we examined dynamical behavior of the SU($n$) spin model at the mean-field level, finding that long-time observables obey simple scaling relations with $n$, and that when $n>2$ dynamical behavior can be highly sensitive to intra-spin coherences.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Three-laser drive with complex amplitudes}
+\section{Relaxing assumptions of the three-laser drive}
 \label{sec:full_drive}
 
-In order to arrive at the single-spin drive Hamiltonian in Eq.~\eqref{eq:drive_single} of the main text, we made the simplifying assumption that all drive amplitudes were real, which essentially enforces a phase-locking condition between the driving lasers.
-If the drive amplitudes are complex, $\Omega_m\to\Omega_me^{-\i\eta_m}$ (with real $\Omega_m,\eta_m$), then the single-spin drive Hamiltonian becomes
+In order to arrive at the drive Hamiltonian in Eq.~\eqref{eq:H_3LD_single} of the main text, we made two simplifying assumptions:
+\begin{enumerate*}
+\item that the excited-state hyperfine manifold had the same total spin $s$ as the ground-state manifold, and
+\item that all drive amplitudes are real, which enforces a phase-locking condition between the driving lasers.
+\end{enumerate*}
+If the excited-state hyperfine manifold has total spin $s+r$ (with $r\in\set{0,+1,-1}$), then the spin-raising/lowering operators $\s_\pm$ that appear in Eq.~\eqref{eq:H_drive_full} (where $\s_{\v v,\pm,j}$ raises or lowers the spin of atom $j$ along axis $\v v$) become
+\begin{align}
+  \s_\pm^{(r)}
+  \equiv -\sqrt{\f{n(n+1)(n-1)}{6}} \times \hat T_\pm^{(r)},
+  &&
+  \hat T_\pm^{(r)} \equiv \mp\sqrt{\f{2(s+r)+1}{2\ell+1}}
+  \sum_\mu \bk{s\mu;1,\pm1|s+r,\mu\pm1} \op{\mu\pm1}{\mu},
+\end{align}
+where $\bk{j_1m_1;j_2m_2|j_3m_3}$ is a Clebsch-Gordan coefficient, and we have normalized $\hat T_\pm^{(r)}$ such that $\tr\sp{\hat T_\pm^{(r)}{}^\dag \hat T_\pm^{(r)}}=1$.
+Still assuming real drive amplitudes, the corresponding effective drive Hamiltonian that replaces Eq.~\eqref{eq:H_3LD_single} is then
+\begin{align}
+  \H_{\t{3LD}}^{\t{single}}
+  = f_r^{(1)} \sp{\tilde\Omega_+ \tilde\Omega_- \s_\z
+    + \tilde\Omega_0 \tilde\Omega_- \s_\x}
+  + f_r^{(2)} \sp{\tilde\Omega_0 \tilde\Omega_+
+    (\s_\z \s_\x  + \s_\x \s_\z)
+    - \p{\tilde\Omega_0^2 \s_\z^2
+      + \tilde\Omega_+^2 \s_\x^2
+      + \tilde\Omega_-^2 \s_\y^2}}
+  - f_r^{(3)} \sum_m \tilde\Omega_m^2,
+\end{align}
+where $f_r^{(k)}$ are $n$-dependent scalars:
+\begin{align}
+  f_0^{(1)} &= 1, &
+  f_{+1}^{(1)} &= -s, &
+  f_{-1}^{(1)} &= s+1, \\
+  f_0^{(2)} &= 1, &
+  f_{+1}^{(2)} &= -\f{s}{n+2}, &
+  f_{-1}^{(2)} &= -\f{s+1}{n-2}, \\
+  f_0^{(3)} &= 0, &
+  f_{+1}^{(3)} &= \f{s(s+1)^2}{n+2}, &
+  f_{-1}^{(3)} &= \f{s^2(s+1)}{n-2}.
+\end{align}
+If additionally the drive amplitudes are complex, $\Omega_m\to\Omega_me^{-\i\eta_m}$ (with real $\Omega_m,\eta_m$), then
 \begin{multline}
-  \Delta \times \H_{\t{3LD}}^{\t{single}}
-  = \p{\Omega_+^2 - \Omega_-^2} \s_\z
-  - \sum_{\sigma\in\set{\pm1}} \Omega_0 \Omega_\sigma
-  \p{\sigma \s_{\tilde\eta_\sigma,\x}
-    + \s_\z \s_{\tilde\eta_\sigma,\x}
-    + \s_{\tilde\eta_\sigma,\x} \s_\z} \\
-  - \Omega_0^2 \s_\z^2 - \p{\Omega_+ + \Omega_-}^2 \s_{\tilde\eta_0,\x}^2
-  - \p{\Omega_+ - \Omega_-}^2 \s_{\tilde\eta_0,\y}^2,
+  \H_{\t{3LD}}^{\t{single}}
+  = f_r^{(1)} \tilde\Omega_+ \tilde\Omega_- \s_\z
+  + \tilde\Omega_0 \sum_{\sigma\in\set{\pm1}}
+  \f{\tilde\Omega_+ + \sigma\tilde\Omega_-}{2}
+  \sp{f_r^{(1)} \sigma \s_{\tilde\eta_\sigma,\x}
+    + f_r^{(2)} \p{\s_\z \s_{\tilde\eta_\sigma,\x}
+      + \s_{\tilde\eta_\sigma,\x} \s_\z}} \\
+  - f_r^{(2)} \sp{\tilde\Omega_0^2 \s_\z^2
+    + \tilde\Omega_+ \s_{\tilde\eta_0,\x}^2
+    + \tilde\Omega_- \s_{\tilde\eta_0,\y}^2}
+  - f_r^{(3)} \sum_m \tilde\Omega_m^2,
 \end{multline}
-where $\s_{\eta\alpha} \equiv e^{-\i\eta \s_\z} \s_\alpha e^{\i\eta \s_\z}$ is a rotated spin-$\alpha$ operator (with e.g.~$\s_{\pi/2,\x}=\s_\y$), and
+where $\s_{\eta\alpha} \equiv e^{-\i\eta \s_\z} \s_\alpha e^{\i\eta \s_\z}$ is a rotated spin-$\alpha$ operator (e.g.~$\s_{\pi/2,\x}=\s_\y$), and
 \begin{align}
@@
   &&
-  \tilde\eta_0 \equiv \f{\eta_+ - \eta_-}{2}.
-\end{align}
+  \tilde\eta_0 \equiv \f{\eta_+ - \eta_-}{2},
+\end{align}
+are the relative phases of the drive amplitudes.
 
@@
   = \f1{12} (n+1) n (n-1)
-  \equiv \eta^2.
-\end{align}
-The Lax formulation (following Refs.~\cite{yuzbashyan2005nonequilibrium, yuzbashyan2006dynamical, yuzbashyan2006relaxation, yuzbashyan2015quantum, smale2019observation}) requires all single-body operators involved to have the same normalization, so we substitute $\s_{\tilde\z,q}\equiv \s_{\z,q}/\eta$ to expand
+  \equiv \xi^2.
+\end{align}
+The Lax formulation (following Refs.~\cite{yuzbashyan2005nonequilibrium, yuzbashyan2006dynamical, yuzbashyan2006relaxation, yuzbashyan2015quantum, smale2019observation}) requires all single-body operators involved to have the same normalization, so we substitute $\s_{\tilde\z,q}\equiv \s_{\z,q}/\xi$ to expand
 \begin{align}
@@
   = -\f1{2N} \sum_{\mu,\nu} \S_{\mu\nu} \S_{\nu\mu}
-  + \eta h \sum_q \sin\p{q} \s_{\tilde\z,q},
+  + \xi h \sum_q \sin\p{q} \s_{\tilde\z,q},
   &&
@@
   = \f1N \sum_q \f{\s_{\alpha,q}}{z-\sin q}
-  + \delta_{\alpha,\tilde\z} \, \eta h,
+  + \delta_{\alpha,\tilde\z} \, \xi h,
 \end{align}
@@
   = \I\p{z} \bar s_\alpha
-  + \delta_{\alpha,\tilde\z} \, \eta h,
+  + \delta_{\alpha,\tilde\z} \, \xi h,
 \end{align}
@@
   = \I\p{z}^2 \sum_{\alpha\ne\tilde\z} \bar s_\alpha^2
-  + \sp{\I\p{z} \bar s_{\tilde\z} + \eta h}^2,
+  + \sp{\I\p{z} \bar s_{\tilde\z} + \xi h}^2,
 \end{align}
@@
   = \I\p{z}^2 \p{Q^2 - \bar s_{\tilde\z}^2}
-  + \sp{\I\p{z} \bar s_{\tilde\z} + \eta h}^2
-  = \I\p{z}^2 Q^2 + \eta^2 h^2
-  + 2 \I\p{z} \eta h \bar s_{\tilde\z}.
+  + \sp{\I\p{z} \bar s_{\tilde\z} + \xi h}^2
+  = \I\p{z}^2 Q^2 + \xi^2 h^2
+  + 2 \I\p{z} \xi h \bar s_{\tilde\z}.
 \end{align}
@@
 \begin{align}
-  \vec\ell\p{z}^2 = \f{Q^2}{z^2-1} + \eta^2 h^2,
+  \vec\ell\p{z}^2 = \f{Q^2}{z^2-1} + \xi^2 h^2,
 \end{align}
@@
 \begin{align}
-  z = \pm \sqrt{1 - \p{\f{Q}{\eta h}}^2}.
+  z = \pm \sqrt{1 - \p{\f{Q}{\xi h}}^2}.
   \label{eq:lax_zeros}
@@
 \begin{align}
-  h_\crit^2 \stackrel{?}{=} \f{Q^2}{\eta^2},
+  h_\crit^2 \stackrel{?}{=} \f{Q^2}{\xi^2},
 \end{align}
@@
 \begin{align}
-  h_\crit^2 \stackrel{?}{=} \f{n-1}{n} \times \f{12}{(n+1)n(n-1)}
+  h_\crit^2 \stackrel{?}{=} \f{n-1}{n} \times \f{12}{n(n+1)(n-1)}
   = \f{12}{n^2\p{n+1}}.

### 2021-04-26 sun_phases/spin_model_paper/figures/make_plots.py

@@ -317,7 +317,7 @@ axis.annotate(r"$\ket\downarrow$", (q_i-0.1,E_i+0.1), va = "bottom", ha = "right
 axis.annotate(r"$\ket\uparrow$", (q_f-0.1,E_f+0.1), va = "bottom", ha = "right")
 
 ##################################################
-# SU(2), gauged frame
+# SU(2), gauge frame
 
 axis = axes[1,0]
 axis.plot(momenta, -np.cos(momenta-phi/2), color = dn_color)
@@ -376,7 +376,7 @@ axis.annotate(r"$\ket{\mu+1}$", (q_2+0.6,E_2), ha = "left", va = "center")
 axis.annotate(r"$\ket{\mu+2}$", (q_3-0.1,E_3), ha = "right", va = "bottom")
 
 ##################################################
-# SU(d), gauged frame
+# SU(d), gauge frame
 
 phi = np.pi/2 * 0.8
 qq = np.pi*5/9
@@ -427,7 +427,7 @@ axes[0,0].set_title("$n=2$")
 axes[0,1].set_title(f"$n={levels}$")
 
 for loc, tag in [ [ (0,1), "lab" ],
-                  [ (1,1), "gauged" ] ]:
+                  [ (1,1), "gauge" ] ]:
     axis = axes[loc].twinx()
     axis.yaxis.tick_right()
     axis.set_yticks([])

### 2021-04-28 sun_phases/spin_model_paper/main.tex
last major updates before blue-forming the SU(n) spin model paper

@@
 
-\title{Engineering an infinite-range SU($n$) spin model with ultracold fermions in an optical lattice}
+\title{Engineering infinite-range SU($n$) interactions with spin-orbit coupled fermions in an optical lattice}
 \author{Michael A. Perlin}
@@
 \email{mika.perlin@gmail.com}
-\author{Diego Barbarena}
+\author{Diego Barberena}
 \affiliation{\JILA}
@@
 \begin{abstract}
-We study multilevel fermions in an optical lattice, with dynamics described by the Hubbard model with on site SU($n$)-symmetric interactions.
-We show that in an appropriate parameter regime this system can be mapped onto a multilevel spin model with all-to-all SU($n$)-symmetric couplings.
+We study multilevel fermions in an optical lattice described by the Hubbard model with on site SU($n$)-symmetric interactions.
+We show that in an appropriate parameter regime this system can be mapped onto a spin model with all-to-all SU($n$)-symmetric couplings.
 Raman pulses that address internal spin states modify the atomic dispersion relation and induce spin-orbit coupling, which can act as a synthetic inhomogeneous magnetic field that competes with the SU($n$) exchange interactions.
@@
 The extension of SU($2$) Hubbard and spin models to SU($n$) has led to predictions of exotic phases of matter such as valence bond solids \cite{read1989valencebond, rokhsar1990quadratic, kaul2012lattice, hermele2011topological} and chiral spin liquids \cite{hermele2009mott, hermele2011topological, chen2016syntheticgaugefield, nataf2016chiral}, as well as the potential to perform universal topological quantum computation \cite{freedman2004class, nayak2008nonabelian} and other phenomena \cite{nataf2014exact, nataf2016exact}.
-Furthermore, disordered SU($n$) spin models have opened analytically tractable avenues for studying quantum chaos and information scrambling \cite{sachdev1993gapless, bentsen2019integrable}.
+Furthermore, disordered SU($n$) spin models have opened analytically tractable avenues for studying quantum chaos and information scrambling \cite{sachdev1993gapless}.
 
@@
 As a result, AMO experiments can directly probe the role of SU($n$) interactions in controllable settings.
-Recent progress includes the experimental observations of bosonization and faster thermalization in multi-component Fermi gas with SU($n$) symmetry \cite{hazzard2012hightemperature, messio2012entropy, bonnes2012adiabatic, stellmer2013degenerate, yip2014theory, pagano2014onedimensional, choudhury2020collective, song2020evidence, sonderhouse2020thermodynamics}, SU($n$) Hubbard phases and phase transitions \cite{taie2012su, hofrichter2016direct, taie2020observation}, two-orbital SU($n$) magnetism \cite{cappellini2014direct, scazza2014observation, zhang2014spectroscopic, beverland2016realizing}, and multi-body SU($n$)-symmetric interactions \cite{goban2018emergence, perlin2019effective}.
+Recent progress includes studies of the thermodynamic properties of SU($n$) fermionic gases \cite{hazzard2012hightemperature, bonnes2012adiabatic, stellmer2013degenerate, yip2014theory, pagano2014onedimensional, choudhury2020collective, song2020evidence, sonderhouse2020thermodynamics}, SU($n$) Hubbard phases and phase transitions \cite{taie2012su, hofrichter2016direct, taie2020observation}, single- \cite{messio2012entropy} and two-orbital \cite{cappellini2014direct, scazza2014observation, zhang2014spectroscopic, beverland2016realizing} SU($n$) magnetism, and multi-body SU($n$)-symmetric interactions \cite{goban2018emergence, perlin2019effective}.
 
@@
 
-Finally, we take a moment to discuss individual $n$-level spins.
-The state of a two-level spin, or a qubit, is commonly represented by a point on (or within) the Bloch sphere.
-More generally, the state $\ket\psi$ of a $n$-level spin can be represented by a probability distribution\footnote{Strictly speaking, in order to make $Q_\psi$ (and $Q_{\hat\rho}$) a probability distribution, one has to divide it by a normalization factor $\frac{4\pi}{d}$.
-We omit this normalization factor for convenience.} $Q_\psi$ on the sphere $\SS^2$.
-The value $Q_\psi\p{\v v}$ at a point $\v v\in\SS^2$ equal to the overlap of $\ket\psi$ with a pure state $\ket{\v v}$ maximally polarized along $\v v$: $Q_\psi\equiv\abs{\bk{\v v|\psi}}^2$ (see Figure \ref{fig:spin_dist}).
-In the case of a mixed state $\hat\rho$, this distribution is defined by $Q_{\hat\rho}\equiv\bk{\v v|\hat\rho|\v v}$\footnote{The dimension $n$ is reflected in $Q_{\hat\rho}$ by the fact that it can be decomposed into spherical harmonics $Y_{\ell m}$ with degree $\ell<n$ (that is, $Q_{\hat\rho}$ is ``low resolution'' or ``band-limited'' at degree $n$) \cite{perlin2020qudit}, as well as in the normalization $\int_{\v v\in\SS^2}\dd\v v\,Q_{\hat\rho}\p{\v v}=\frac{4\pi}{d}$.}.
-Closely related spherical representations of multilevel spin states (and operators) are discussed in Ref.~\cite{li2013weylwignermoyal}.
-
 \begin{figure}
@@
 \caption{
-Whereas the state of a two-level spin (qubit) can be represented by a point on (or inside) the Bloch sphere, the state of a $n$-level spin is more generally represented by a probability distribution on the sphere, decomposable into spherical harmonics $Y_{\ell m}$ of degree $\ell<n$.
+Whereas the state of a two-level spin (qubit) can be represented by a point on (or inside) the Bloch sphere, the state of a $n$-level spin is more generally represented by a probability distribution on the sphere.
 The distribution shown for $n=10$ corresponds to a Haar-random pure state.
@@
 
-In fact, $Q_{\hat\rho}$ is essentially the Husimi-$Q$ function commonly used (e.g.~in the spin squeezing community \cite{ma2011quantum}) to represent collective states of spin-$\frac{1}{2}$ particles in the Dicke manifold.
-This correspondence is made precise by identifying the Hilbert space of a single $n$-level spin with the Dicke manifold of $n-1$ spin-$\frac{1}{2}$ particles.
-The Hilbert space of a 10-level (spin-$\frac92$) nucleus of a ${}^{87}$Sr atom, for example, can be thought of as the Dicke manifold of 9 spin-$\frac{1}{2}$ particles (``excess'' protons and neutrons).
-We will freely use this correspondence to import standard results and constructions for the Dicke manifold (e.g.~the spin operators $\s_\x,\s_\y,\s_\z$) into the context of a multilevel spin.
+Finally, we take a moment to discuss individual $n$-level spins.
+The state of a two-level spin, or a qubit, is commonly represented by a point on (or within) the Bloch sphere.
+More generally, the state $\ket\psi$ of a $n$-level spin can be represented by a quasi-probability distribution $Q_\psi$ on the sphere (commonly known as the Husimi-$Q$ function, e.g.~in the spin-squeezing community \cite{ma2011quantum}).
+The value $Q_\psi\p{\v v}$ at a point $\v v$ on the sphere is equal to the overlap of $\ket\psi$ with a pure state $\ket{\v v}$ maximally polarized in the direction of $\v v$: $Q_\psi\equiv\abs{\bk{\v v|\psi}}^2$ (see Figure \ref{fig:spin_dist}).
+In the case of a mixed state $\hat\rho$, this distribution is defined by $Q_{\hat\rho}\equiv\bk{\v v|\hat\rho|\v v}$.
+Closely related spherical representations of multilevel spin states and operators are discussed in Refs.~\cite{dowling1994wigner, li2013weylwignermoyal}.
+In practice, it is conceptually useful to identify the Hilbert space of a single $n$-level spin with the Dicke manifold of $n-1$ spin-$\frac12$ particles.
 
@@
 \end{align}
-where $m\in\set{+1,0,-1}$ indexes the laser pointing along $\v v_m$; and $\s_{0,j}\equiv \s_{\z,j}$ for shorthand.
+where $m\in\set{+1,0,-1}$ indexes the laser pointing along $\v v_m$; $\s_{0,j}\equiv\s_{\z,j}$ for shorthand; and $\s_{\z,j},\s_{+,j}\s_{-,j}$ are standard axial, spin-raising, and spin-lowering operators for spin $j$.
 In the far-detuned limit $\abs{\Delta}\gg\abs{\Omega_m}$, a second-order perturbative treatment of electronic excitations ($\ket\e$) yields an effective drive Hamiltonian that only addresses ground-state nuclear spins.
@@
   e^{-\i\frac{\pi}{2}\p{\s_\y + \s_\y^2}} \ket{s}
-  \propto \ket{s} + \ket{-s}.
+  \stackrel{n\,\t{even}}{\propto} \ket{s} + \ket{-s}.
 \end{align}
@@
 The two-level SOC drive in Eq.~\eqref{eq:drive_2} has been implemented with an external laser that couples the two electronic states of nuclear-spin-polarized atoms, with $\dn$ ($\up$) indexing the ground (excited) electronic state \cite{wall2016synthetic, livi2016synthetic, kolkowitz2016spinorbitcoupled, bromley2018dynamics, he2019engineering}.
-In contrast, the drive we considered in Section \ref{sec:controls} addresses electronic excitations off-resonantly, inducing an effective Hamiltonian in the ground-state hyperfine manifold with spin projections $\mu\in\set{s,s-1,\cdots,-s}$ (a similar one-laser scheme was used to study SOC using a subspace of the ground-state manifold in Ref.~\cite{mancini2015observation}).
+In contrast, the drive we considered in Section \ref{sec:controls} addresses electronic excitations off-resonantly, inducing an effective Hamiltonian in the ground-state hyperfine manifold with spin projections $\mu\in\set{s,s-1,\cdots,-s}$ (a similar scheme was used to study SOC in a subspace of the ground-state manifold in Ref.~\cite{mancini2015observation}).
 Nonetheless, both the two-level drive in Eq.~\eqref{eq:drive_2} and the $n$-level drive in Eq.~\eqref{eq:H_3LD} become homogeneous (i.e.~independent of the spatial mode index $j$ or $q$) and independent of the SOC angle $\phi$ after the same spin-symmetric gauge transformation\footnote{Note that the ``asymmetric'' gauge transformation $(\c_{j,\up}^\dag,\c_{j,\dn}^\dag)\to(e^{\i\phi j} \c_{j,\up}^\dag,\c_{j,\dn}^\dag)$, sometimes performed in the two-state SOC literature, does not generalize as nicely to $n>2$.} $\c_{j\mu}^\dag\to e^{\i\phi\mu j} \c_{j\mu}^\dag$.
@@
   = -2J \sum_q
-  \sp{\cos\p{q} \hat w_{+,q}^{(\phi)} - \sin\p{q} \hat w_{+,q}^{(\phi)}},
+  \sp{\cos\p{q} \hat w_{+,q}^{(\phi)} - \sin\p{q} \hat w_{-,q}^{(\phi)}},
 \end{align}
@@
 Colors indicate different spin projections $\mu$.
-In the ``lab frame'' ({\bf a},{\bf b}), kinetic energy is insensitive to spin, but the drive imparts a momentum kick $q\to q+\phi$ into any atom that absorbs a photon.
-Changing to the ``gauge frame'' ({\bf c},{\bf d}), essentially by shifting the momentum label $q$ for each spin state $\mu$, makes the drive diagonal in the momentum index, but comes at the cost of making kinetic energy spin-dependent.
+In the ``lab frame'' ({\bf a},{\bf b}), kinetic energy is insensitive to spin, but a spin transition $\mu\to\mu+1$ is accompanied by a momentum kick $q\to q+\phi$ from the drive.
+Changing into the ``gauge frame'' ({\bf c},{\bf d}), essentially by shifting the momentum label $q$ for each spin state $\mu$, makes the drive diagonal in the momentum index, but comes at the cost of making kinetic energy spin-dependent.
 }
@@
 We now study the dynamical behavior of the SOC spin Hamiltonian $H_\spin$ in Eq.~\eqref{eq:H_spin}, and henceforth work exclusively in the ``gauge frame'' of $\H_\spin$ and the three-laser drive $\H_{\t{3LD}}$ in Eq.~\eqref{eq:H_3LD}.
-The general idea will be a Ramsey-type setup in which we first prepare an initial state with the three-laser drive (using fast pule sequences), let this state evolve freely for some time under $\H_\spin$, and then again use the three-laser drive to map observables of interest onto spin projection measurements (e.g.~with spin qudit tomography \cite{newton1968measurability, perlin2020qudit}).
-At the mean-field (MF) level, up to constant energy shifts the undriven spin Hamiltonian becomes
+We use a Ramsey-like setup wherein we prepare an initial state with the three-laser drive (using fast pule sequences), then let the state evolve freely for some time under $\H_\spin$, and finally apply again the three-laser drive to map observables of interest onto spin projection measurements (e.g.~with spin qudit tomography \cite{newton1968measurability, perlin2020qudit}).
+At the mean-field (MF) level, the undriven spin Hamiltonian (neglecting constant energy shifts) becomes
 \begin{align}
@@
 This decomposition requires no approximations, and reduces the number of variables to keep track of by a factor of $\sim n$.
-See Appendices \ref{sec:MFT} and \ref{sec:bosons} additional details about our numerical simulations and the Schwinger boson equations of motion.
+See Appendices \ref{sec:MFT} and \ref{sec:bosons} for additional details about our numerical simulations and the Schwinger boson equations of motion.
 
@@
 The time-averaged transverse magnetization $\sigma_\MF$ then trivially vanishes as $h\to\infty$.
-The interaction energy $\bk{\ss}_\MF = \bk{\bar{\v s}}_\MF \cdot \bk{\bar{\v s}}_\MF$, meanwhile, has contributions from:
+The interaction energy $\bk{\ss}_\MF = \bk{\bar{\v s}}_\MF \cdot \bk{\bar{\v s}}_\MF + O(1/N)$, meanwhile, has contributions from:
 \begin{enumerate*}
@@
 The weak-field effective Hamiltonian preserves permutational symmetry, so $\bbk{\ss}_\MF\to1$ as $h\to0$.
-Moreover, the initial y-magnetization $\bk{\S_\y}=0$ is conserved by $\H_\spin$, so long-time-averaged magnetization $\sigma_\MF$ is determined by the time-average of $\hat\sigma_\x\equiv\s_\x/s$ for a single spin:
+Moreover, the initial y-magnetization $\bk{\S_\y}=0$ is conserved by $\H_\spin$, so long-time-averaged magnetization $\sigma_\MF$ is determined by the time-average of $\s_\x$ for a single spin:
 \begin{align}
   \lim_{h\to0} \sigma_\MF
-  = \abs{\lim_{T\to\infty} \f1T \int_0^T \dd\tau
-  \bk{\x|\hat\sigma_\x\p{\tau}|\x}},
+  = \f1s \abs{\lim_{T\to\infty} \f1T \int_0^T \dd\tau
+  \bk{\x|\s_\x\p{\tau}|\x}},
 \end{align}
@@
 \begin{align}
-  \hat\sigma_\x\p{\tau} = e^{\i\tau \s_\z^2} \hat\sigma_\x e^{-\i\tau \s_\z^2}.
-\end{align}
-By mapping the Hilbert space of an $n$-level spin onto the PS (Dicke) manifold of $n-1$ spin-$\frac12$ particles, we can adapt exact analytical results for the Ising model \cite{foss-feig2013nonequilibrium}\footnote{See Appendix K of Ref.~\cite{perlin2020shorttime} for a simpler adaptation of the analytics in Ref.~\cite{foss-feig2013nonequilibrium} to the one-axis twisting model $\H_{\t{OAT}}=\chi \s_\z^2$.} to find that
-\begin{align}
-  \bk{\x|\hat\sigma_\x\p{\tau}|\x} = \p{\cos\tau}^{n-2},
+  \s_\x\p{\tau} = e^{\i\tau \s_\z^2} \s_\x e^{-\i\tau \s_\z^2}.
+\end{align}
+We can adapt exact analytical results for the dynamics of an infinite-range Ising model \cite{foss-feig2013nonequilibrium}\footnote{See Appendix K of Ref.~\cite{perlin2020shorttime} for a simpler adaptation of the analytics in Ref.~\cite{foss-feig2013nonequilibrium} to the one-axis twisting model $\H_{\t{OAT}}=\chi \s_\z^2$.} to find that
+\begin{align}
+  \bk{\x|\s_\x\p{\tau}|\x} = s \p{\cos\tau}^{n-2},
 \end{align}
@@
 \end{align}
-When going beyond mean-field theory, inter-spin correlations generated by $\S_\z^2$ in Eq.~\eqref{eq:H_spin_eff} will cause $\bk{\S_\x}$ (and thereby the magnetization $\bk{\vec\sigma}$) to decay on time scales that diverge as $N\to\infty$.
+When going beyond mean-field theory, inter-spin correlations generated by $\S_\z^2$ in Eq.~\eqref{eq:H_spin_eff} will cause $\bk{\S_\x}$ (and thereby the magnetization $\bk{\vec\sigma}$) to decay as $e^{-O(t^2/Ns)}$; the timescale of this decay diverges as $N\to\infty$.
 On a lattice of linear size $L$ without periodic boundary conditions, additional corrections to the behavior predicted above will appear on $O(L/J)$ timescales.
@@
 \begin{align}
-  \ket\XXI \simeq \T_\z^{\pi/2} \R_\z^{\pi/2} \ket\X,
+  \ket\XXI \simeq \R_\z^{\pi/2} \T_\z^{\pi/2} \ket\X,
 \end{align}
@@
   e^{-\i t \H_\spin} \ket\XXI
-  &\simeq e^{-\i t \H_\spin} \T_\z^{\pi/2} \R_\z^{\pi/2} \ket\X \\
-  &\simeq \T_\z^{\pi/2} \R_\z^{\pi/2} e^{-\i t \H_\spin} \ket\X.
+  &\simeq e^{-\i t \H_\spin} \R_\z^{\pi/2} \T_\z^{\pi/2} \ket\X \\
+  &\simeq \R_\z^{\pi/2} \T_\z^{\pi/2} e^{-\i t \H_\spin} \ket\X.
 \end{align}
@@
 \begin{align}
-  \R_\z^\theta{}^\dag \T_\z^\theta{}^\dag \, \ss \, \T_\z^\theta \R_\z^\theta
+  \T_\z^\theta{}^\dag \R_\z^\theta{}^\dag \, \ss \, \R_\z^\theta \T_\z^\theta
   = \ss,
@@
   \sigma_\MF = \abs{\bbk{\vec\sigma}_\MF}
-  = \abs{\bbk{\R_\z^\theta{}^\dag \T_\z^\theta{}^\dag \,
-      \vec\sigma \, \T_\z^\theta \R_\z^\theta}_\MF}
+  = \abs{\bbk{\T_\z^\theta{}^\dag \R_\z^\theta{}^\dag \,
+      \vec\sigma \, \R_\z^\theta \T_\z^\theta}_\MF}
 \end{align}
@@
 We then introduced external control fields, finding a simple three-laser drive that homogeneously addresses nuclear spins with a variety of spin Hamiltonians.
-Taking a closer look at the effect of the spin-orbit coupling (SOC) induced by the driving lasers, we found that maintaining integrity of the spin model requires weak SOC, which in turn gives rise to a (synthetic) inhomogeneous magnetic field.
+Taking a closer look at the effect of the spin-orbit coupling (SOC) induced by the driving lasers, we found that maintaining the validity of the spin model requires weak SOC, which in turn gives rise to a (synthetic) inhomogeneous magnetic field.
 Finally, we examined dynamical behavior of the SU($n$) spin model at the mean-field level, finding that long-time observables obey simple scaling relations with $n$, and that when $n>2$ dynamical behavior can be highly sensitive to intra-spin coherences.
@@
 Our work makes important progress in understanding the SU($n$) Fermi-Hubbard model in experimentally relevant parameter regimes, and we expect our findings to be readily testable in experiments with ultracold atoms.
-Given the possibility for long-range SU($n$) interactions, we hope our work stimulates further efforts into simulating SY and SYK-like models \cite{sachdev1993gapless} in cold atomic platforms.
+Given the possibility for long-range SU($n$) interactions, we hope our work stimulates further efforts into simulating SY and SYK-like models \cite{sachdev1993gapless, bentsen2019integrable} in cold atomic platforms.
 In follow-up work, it would be interesting to study the the relationship between initial states and dynamical phases of our SU($n$) spin model more systematically, and to consider the effect of quantum corrections to mean-field behavior.
@@
 \end{enumerate*}
-If the excited-state hyperfine manifold has total spin $s+r$ (with $r\in\set{0,+1,-1}$), then the spin-raising/lowering operators $\s_\pm$ that appear in Eq.~\eqref{eq:H_drive_full} (where $\s_{\v v,\pm,j}$ raises or lowers the spin of atom $j$ along axis $\v v$) become
+If the excited-state hyperfine manifold has total spin $s+r$ (with $r\in\set{+1,0,-1}$), then the spin-raising/lowering operators $\s_\pm$ that appear in Eq.~\eqref{eq:H_drive_full} (where $\s_{\v v,\pm,j}$ raises or lowers the spin of atom $j$ along axis $\v v$) become
 \begin{align}

### 2021-07-28 sun_phases/spin_qudit_tomo.tex
updated tomography paper to reflect/discuss relevant literature just a few important todos remain

@@
-\documentclass[a4paper,twocolumn,unpublished]{quantumarticle}
-\pdfoutput=1
+\documentclass[nofootinbib,twocolumn]{revtex4-1}
 
@@
 
-%%% bibliography
-\usepackage[backend=bibtex,style=phys,doi=false]{biblatex}
-\addbibresource{\jobname.bib}
-
 %%% linking references
+\usepackage[dvipsnames]{xcolor}
 \usepackage{hyperref}
@@
 \title{Spin qudit tomography}
-\author[]{Michael A.~Perlin}
+\author{Michael A.~Perlin}
 \email{mika.perlin@gmail.com}
-\orcid{https://orcid.org/0000-0002-9316-1596}
-\author[]{Diego Barberena}
-\author[]{Ana Maria Rey}
-\orcid{http://orcid.org/0000-0001-7176-9413}
-\affiliation[1]{JILA, National Institute of Standards and Technology and University of Colorado, 440 UCB, Boulder, Colorado 80309, USA}
-\affiliation[2]{Center for Theory of Quantum Matter, University of Colorado, Boulder, CO, 80309, USA}
+\author{Diego Barberena}
+\author{Ana Maria Rey}
+\affiliation{JILA, National Institute of Standards and Technology and University of Colorado, 440 UCB, Boulder, Colorado 80309, USA}
+\affiliation{Center for Theory of Quantum Matter, University of Colorado, Boulder, CO, 80309, USA}
 
@@
 \begin{abstract}
-  We consider the task of performing quantum state tomography on a $d$-state spin qudit, using only measurements of spin projection onto different quantization axes.
-  By an exact mapping onto the classical problem of signal recovery on the sphere, we prove that full reconstruction of arbitrary qudit states requires a minimal number of measurement axes, $r_d^{\t{min}}$, that is bounded by $2d-1\le r_d^{\t{min}}\le d^2$.
-  We conjecture that $r_d^{\t{min}}=2d-1$, which we verify numerically for all $d\le200$.
+We consider the task of performing quantum state tomography on a $d$-level spin qudit, using only measurements of spin projection onto different quantization axes.
+After introducing a basis of operators closely related to the spherical harmonics, which obey the rotational symmetries of spin qudits, we map our quantum tomography task onto the classical problem of signal recovery on the sphere.
 We then provide algorithms with $O\p{rd^3}$ serial runtime, parallelizable down to $O\p{rd^2}$, for (i) computing a priori upper bounds on the expected error with which spin projection measurements along $r$ given axes can reconstruct an unknown qudit state, and (ii) estimating a posteriori the statistical error in a reconstructed state.
@@
 Quantum state tomography, the task of reconstructing a quantum state by collecting and processing measurement data, is an essential primitive for quantum sensing, quantum simulation, and quantum information processing.
-The central importance of quantum state tomography has led to the development of techniques based on least-squares inversion \cite{opatrny1997leastsquares}, linear regression \cite{qi2013quantum}, maximum likelihood estimation \cite{teo2011quantumstate, smolin2012efficient}, Bayesian inference \cite{huszar2012adaptive, ferrie2014quantum, granade2016practical}, compressed sensing \cite{gross2010quantum, kalev2015quantum}, neural networks \cite{torlai2018neuralnetwork}, among others too numerous to exhaustively list here.
-These techniques are typically developed in a general information-theoretic setting, and make minimal assumptions about the physical medium of a quantum state.
+The central importance of quantum state tomography has led to the development of techniques based on least-squares inversion \cite{opatrny1997leastsquares}, linear regression \cite{qi2013quantum}, maximum likelihood estimation \cite{teo2011quantumstate, smolin2012efficient}, Bayesian inference \cite{huszar2012adaptive, ferrie2014quantum, granade2016practical}, compressed sensing \cite{gross2010quantum, kalev2015quantum}, and neural networks \cite{torlai2018neuralnetwork}, among others.
+These techniques are typically developed in a general, information-theoretic setting, and make minimal assumptions about the physical medium of a quantum state.
 As a consequence, even well-established techniques can be ill-suited for physical platforms with unique or limited capabilities.
@@
 Due to advancements in experimental capabilities to address nuclear spin states (i.e.~hyperfine levels) in ultracold atomic systems \cite{daley2011quantum, mischuck2012control, smith2013quantum, cazalilla2014ultracold}, as well as developments in the control of ultracold molecular systems \cite{bohn2017cold, zeppenfeld2012sisyphus, kozyryev2017sisyphus, puri2017synthesis, wu2017cryofuge, marco2019degenerate, liu2019molecular, anderegg2019optical, chou2020frequencycomb, lin2020quantum}, a particular setting of growing interest is the spin qudit, or a multilevel quantum angular momentum degree of freedom.
-Spin qudits can provide advantages over their qubit counterparts for quantum sensing \cite{hemmer2018squeezing, evrard2019enhanced}, enable quantum simulations of SU($d$) magnetism \cite{cazalilla2014ultracold, banerjee2013atomic, zhang2014spectroscopic, scazza2014observation, goban2018emergence}, and offer unique capabilities for quantum error correction \cite{albert2020robust, gross2020encoding}.
+Spin qudits can provide advantages over their qubit counterparts for quantum sensing \cite{hemmer2018squeezing, evrard2019enhanced}, enable quantum simulations of SU($d$) magnetism \cite{cazalilla2014ultracold, banerjee2013atomic, zhang2014spectroscopic, scazza2014observation, goban2018emergence}, % todo: cite SU(n) spin model paper!
+and offer unique capabilities for quantum error correction \cite{albert2020robust, gross2021designing}.
 In all cases, quantum state tomography is necessary to take full advantage of a spin qudit.
@@
 
-The general problem of qudit tomography is not new, with an extensive literature on a variety of techniques \cite{manko1997spin, schmied2011tomographic, evrard2019enhanced, flammia2005minimal, thew2002qudit, salazar2012quantum, sosa-martinez2017quantum, ha2018minimal, stefano2019set, palici2020oam}.
-However, existing protocols either rely on infinite-dimensional representations of a quantum spin \cite{manko1997spin, schmied2011tomographic, evrard2019enhanced}, or require the capability to perform essentially arbitrary operations on a qudit \cite{thew2002qudit, flammia2005minimal, salazar2012quantum, sosa-martinez2017quantum, ha2018minimal, stefano2019set, palici2020oam}, generally resulting in tomographic protocols that can be highly inefficient or unachievable in practice.
+The problem of qudit tomography is not new, with an extensive literature on a variety of techniques \cite{newton1968measurability, hofmann2004quantumstate, filippov2010inverse, schmied2011tomographic, evrard2019enhanced, flammia2005minimal, thew2002qudit, salazar2012quantum, sosa-martinez2017quantum, ha2018minimal, stefano2019set, palici2020oam, filippov2010inverse, hofmann2004quantumstate, filippov2010inverse}.
+However, most existing protocols either rely on infinite-dimensional representations of a quantum spin \cite{manko1997spin, schmied2011tomographic, evrard2019enhanced}, or require the capability to perform essentially arbitrary operations on a qudit \cite{thew2002qudit, flammia2005minimal, salazar2012quantum, sosa-martinez2017quantum, ha2018minimal, stefano2019set, palici2020oam}, generally resulting in tomographic protocols that can be highly inefficient or unachievable in practice.
 The protocols based on infinite-dimensional representations of a quantum spin have the advantage of reconstructing its state from measurements of spin projection onto different spatial axes, which are generally accessible with any spin qudit.
-Nonetheless, these protocols obfuscate the minimal requirements for performing full state tomography, provide no straightforward error bounds or guarantees of accuracy, and (with the exception of Ref.~\cite{schmied2011tomographic}) extract only a small fraction of the information contained in measurement data.
-
-In this work, we provide an efficient framework for performing spin qudit tomography using only measurements of spin projection onto different spatial axes.
-We first introduce, in Section \ref{sec:polarization_ops}, a set of qudit operators that are closely related to the spherical harmonics, and which play a central role in our work.
-We then map the quantum problem of spin qudit tomography onto the well-studied classical problem of signal recovery on the sphere in Section \ref{sec:signal_recovery}, thereby importing a host of existing literature and mathematical machinery that has been developed for this task \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
-In particular, this mapping allows us to bound the minimum number of measurement axes necessary to perform full tomography on a $d$-level spin qudit, $r_d^{\t{min}}$, by $2d-1\le r_d^{\t{min}}\le d^2$.
+Nonetheless, in addition to their inefficiencies these protocols obfuscate the minimal requirements for performing full state tomography, provide no straightforward error bounds or guarantees of accuracy, and (with the notable exception of Ref.~\cite{schmied2011tomographic}) generally extract only a small fraction of the information contained in measurement data.
+
+In this work, we consider the task of performing spin qubit tomography using only measurements of spin projection onto different spatial axes, and address some of the shortcomings discussed above.
+This sort of qudit tomography was first considered in Ref.~\cite{newton1968measurability}, as well as a few later works \cite{hofmann2004quantumstate, filippov2010inverse, schmied2011tomographic}.
+Specifically, Ref.~\cite{newton1968measurability} provided an explicit protocol for reconstructing a $d$-level spin qudit state from measurements of spin projection along $2d-1$ axes, the minimum number necessary for full tomography of an arbitrary qudit state.
+However, the protocol in Ref.~\cite{newton1968measurability} involves a choice of a single (arbitrary) angle $\theta$, and provides no means for comparing different choices of $\theta$, which may result in wildly different statistical errors (i.e.~precision) in a reconstructed state.
+Other works provide insightful discussions into the problem of spin qudit tomography, but either
+\begin{enumerate*}
+\item require making assumptions about the qudit state in question \cite{schmied2011tomographic},
+\item do not address the question of statistical error \cite{hofmann2004quantumstate}, or
+\item provide a measure of statistical error that is overly conservative and computationally demanding \cite{filippov2010inverse}.
+\end{enumerate*}
+
+In Section \ref{sec:polarization_ops}, we introduce a set of qudit operators that are closely related to the spherical harmonics, and which play a central role in our work.
+We then map the quantum problem of spin qudit tomography onto the classical problem of signal recovery on the sphere in Section \ref{sec:signal_recovery}, thereby providing an interesting new perspective on spin qudit tomography.
 In Section \ref{sec:error} we provide a priori upper bounds and a posteriori estimates of the statistical error in a qudit state reconstructed from measurements of spin projection along a given set of $r$ measurement axes.
 The capability to determine upper bounds on reconstruction error a priori motivates a simple randomized tomography protocol that we outline in Section \ref{sec:protocol}, and for which we numerically find that using more measurement axes yields substantial benefits that plateau after $r\approx3d$.
-To facilitate the use of our protocols, we make all of our codes publicly available at Ref.~\cite{codes}.
-
-% todo: add figure for spin measurements?
 
@@
   \equiv \bk{\ell m|R\p{\bar{\bm\omega}}|\ell n}
+  \label{eq:wigner_D}
 \end{align}
@@
 We clarify the connection between the quantum problem of reconstructing $\rho$ from spin projection measurements and the classical problem of reconstructing $\rho^{\t{PS}}$ from its values $\rho^{\t{PS}}\p{\bm v}$ in the following section.
-In addition to providing a new perspective on the problem of spin qudit tomography, this connection will allow us to place upper bounds on the minimal number of measurement axes $\bm v$ necessary for full recovery of an unknown state.
 
@@
 \label{sec:signal_recovery}
-
-% todo: add figure for signal recovery problem
 
@@
 
-In fact, the problem of finding suitable axes $V$ and coefficients $C_{\ell m}\p{\bm v}$ to satisfy Eq.~\eqref{eq:tomo_recovery} can be mapped onto the well-studied problem of signal recovery on the sphere \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
+\begin{figure}
+  \centering
+  \includegraphics[width=0.4\columnwidth]{sphere_points.pdf}
+  \caption{Signal recovery on the sphere is the problem of reconstructing an unknown function $f$ (red distribution) from its values $f\p{\bm v}$ at specific points $\bm v\in V$ (blue dots) on the sphere.
+    For most choices of $V$, reconstruction of $f$ is possible if there are at least as many points in $V$ as there are degrees of freedom in $f$.}
+  \label{fig:sphere_points}
+\end{figure}
+
+In fact, the problem of finding suitable axes $V$ and coefficients $C_{\ell m}\p{\bm v}$ to satisfy Eq.~\eqref{eq:tomo_recovery} can be mapped onto the well-studied problem of signal recovery on the sphere (see Figure \ref{fig:sphere_points}) \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
 The signal recovery problem can be stated as follows: given a square-integrable function $f$ on the sphere, with the spherical harmonic expansion
@@
 In practice, spin qudits typically have only a modest dimension $d$, which allows for simpler and optimized tomography protocols that are practical despite worse scaling with $d$ (see Section \ref{sec:protocol}).
+A natural avenue to develop better spin qudit tomography protocols would therefore be to build on the existing classical signal recovery algorithms, tailoring them to solve the reduced recovery problem in Eq.~\eqref{eq:tomo_recovery} rather than the full recovery problem in Eq.~\eqref{eq:full_recovery}.
+We leave these developments to future work.
 
@@
 Moreover, for any fixed degree $\ell$, finding solutions to the reduced recovery problem in Eq.~\eqref{eq:tomo_recovery} is equivalent to the recovery of a degree-$\ell$ function $f_\ell = \sum_m f_{\ell m} Y_{\ell m}$, which is provably possible with $\abs{V}=2\ell+1$ samples \cite{freeden2008spherical}.
-In the case of spin qudit tomography, the degree $\ell$ takes a maximal value of $\ell_{\t{max}}\equiv d-1$, so state recovery requires at least as many measurement axes as there are polarization operators with degree $\ell_{\t{max}}$, namely $2\ell_{\t{max}}+1=2d-1$.
-
-Since
-\begin{enumerate*}
-\item there are at most $2d-1$ polarization operators of any given degree, and
-\item measuring spin projection along a single axis provides measurement data for polarization operators of all degrees,
-\end{enumerate*}
-one might expect that $2d-1$ suitably chosen measurement axes would always suffice to perform full tomography of a $d$-level spin qudit.
-In the language of classical signal recovery, we would expect there to exist a set of $2\ell_{\t{max}}+1$ points on the sphere that enable the recovery of any function with fixed degree $\ell\le\ell_{\t{max}}$, given prior knowledge of $\ell$\footnote{We could further conjecture that if {\it any} set of points $V$ enables the recovery of degree-$\ell_{\t{max}}$ functions on the sphere, then $V$ enables the recovery degree-$\ell$ functions with $\ell\le\ell_{\t{max}}$.
-  However, this condition is stronger than necessary to guarantee full recovery of a $d$-level spin qudit state with $2d-1$ measurement axes, which only requires the existence of {\it some} such set of points.}.
-Unfortunately, we have no general proof of this conjecture, which requires a deeper analysis of fixed-degree signal recovery on the sphere.
-Nonetheless, in the following section we numerically verify that $2d-1$ points suffice for all $d\le200$.
+In the case of spin qudit tomography, the degree $\ell$ takes a maximal value of $\ell_{\t{max}}\equiv d-1$, so state recovery requires as many measurement axes as there are polarization operators with degree $\ell_{\t{max}}$, namely $2\ell_{\t{max}}+1=2d-1$.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Tomographic reconstruction error}
+\section{State reconstruction error}
 \label{sec:error}
 
-For the practically minded, the mere existence of solutions to a problem is less interesting than the exposition of a particular solution.
+For the practically minded, proving the existence of solutions to a problem is less interesting than the exposition of a particular solution.
 On a high level, a spin qudit tomography protocol consists of
@@
 \item selecting a set of measurement axes,
-\item collecting measurement data on spin projection onto these axes, and then
-  \label{tomo:measure}
+\item \label{tomo:measure} collecting measurement data on spin projection onto these axes, and then
 \item processing the collected data to reconstruct the state of the spin qudit.
@@
 \end{align}
+Here $D^\ell_{0,m}\p{\bm v}$ is a Wigner rotation matrix element, defined in Eq.~\eqref{eq:wigner_D}.
 As the singular values of $M_V$ are invariant under unitary transformations, it follows that
@@
     These results do not count fixed runtimes to pre-compute quantities that can be recycled for every new choice of $V$ and $\rho$.
-    Dashed lines show fits to a runtime $t=c d^\alpha$ for the 20 largest values of $d$, finding $\alpha\approx 3.8\pm0.1$.
-    These simulations found axes $V$ with $\epsilon_V<d$ for all $d\le200$, thereby verifying for these cases that $2d-1$ measurement axes are sufficient to perform full state tomography.}
+    Dashed lines show fits to a runtime $t=c d^\alpha$ for the 20 largest values of $d$, finding $\alpha\approx 3.8\pm0.1$.}
   \label{fig:times}
@@
 \end{align}
-where the {\it quantum error scale} $\epsilon_V$, so called because the bound in Eq.~\eqref{eq:bound} is set by quantum shot noise, is defined in terms of (half of) the spectral range of $T_{\ell,0}$:
+where the {\it quantum error scale} $\epsilon_V$, so called because the bound in Eq.~\eqref{eq:bound} is set by quantum shot noise, is defined in terms of the spectral range of $T_{\ell,0}$:
 \begin{align}
@@
 \end{align}
-The covariances $\bbk{\epsilon_{\bm v\ell}\epsilon_{\bm w\ell}}$ are generally determined by the sources of measurement error in any given experiment, but will typically satisfy $\bbk{\epsilon_{\bm v\ell}\epsilon_{\bm w\ell}} = \delta_{\bm v\bm w}\bbk{\epsilon_{\bm v\ell}^2}$.
+The covariances $\bbk{\epsilon_{\bm v\ell}\epsilon_{\bm w\ell}}$ are generally determined by the sources of measurement error in any given experiment, but will typically satisfy $\bbk{\epsilon_{\bm v\ell}\epsilon_{\bm w\ell}} = \delta_{\bm v\bm w}\bbk{\epsilon_{\bm v\ell}^2}$ because measurements along $\bm v$ are independent of measurements along $\bm w$.
 If measurement error is limited by shot noise, then (see Appendix \ref{sec:exact})
@@
     Color indicates the qudit dimension $d$.
-    The rapid initial drop in $\tilde\beta(p)$ implies that using more measurement axes can substantially lower the upper bound on reconstruction error provided in Eq.~\eqref{eq:bound}, and that these benefits plateau after $p\approx d$.}
+    The rapid initial drop in $\tilde\beta(p)$ implies that using more measurement axes can substantially lower the upper bound on reconstruction error provided in Eq.~\eqref{eq:bound}, and that these benefits plateau after $p\approx d$.
+    Horizontal reference lines mark the corresponding (normalized) minimal quantum error scales $\min_\theta\epsilon_{V(\theta)}/\tilde\beta(0)$ achievable with the method in Ref.~\cite{newton1968measurability}, which measures along $2d-1$ axes $V(\theta)$ equally spaced at a polar angle $\theta$.}
   \label{fig:axes}
@@
 The reconstruction error bound in Eq.~\eqref{eq:bound} nominally provides a straightforward answer: at a fixed total number of measurements, $N=n\abs{V}$, the number of measurement axes should be chosen to minimize the (squared) reconstruction error $\E_V\p{\rho}^2<\epsilon_V^2/n\propto\epsilon_V^2\abs{V}$.
-We therefore consider the measurement-adjusted error scale
-\begin{align}
-  \beta(p)
-  \equiv \min\set{\epsilon_V:\abs{V}=2d-1+p} \times \sqrt{2d-1+p}
+We therefore consider the measurement-adjusted error scale $\beta(p)$ defined by
+\begin{align}
+  \beta(p)^2
+  \equiv \min_V\Set{\epsilon_V^2\abs{V}:\abs{V}=2d-1+p}
 \end{align}
@@
 Figure \ref{fig:axes} shows the results of such empirical minimization, where we find that $\tilde\beta(p)$ drops substantially with $p$ before plateauing at $p\approx d$, after which there are only minor benefits to using more measurement axes.
-In the interest of reducing experimental complexity as well as the runtime of our tomography protocol, which grow linearly in $\abs{V}$, we therefore conclude that spin qudit tomography should be performed with $\abs{V}\approx3d$ measurement axes.
+In the interest of reducing experimental complexity as well as the runtime of our randomized tomography protocol, which grow linearly in $\abs{V}$, we therefore conclude that this protocol should be performed with $\abs{V}\approx3d$ measurement axes.
+
+% todo: paragraph on old method's appearance in fig:axes
+
+% todo: appendix on optimizing the old method that appears in fig:axes
 
@@
 
-\printbibliography
-\onecolumn
+\bibliography{spin_qudit_tomo.bib}
+\onecolumngrid
 \newpage

### 2021-07-28 sun_phases/scripts/plot_tomo_axes.py

@@ -16,8 +16,15 @@ plt.rcParams.update(params)
 
 figsize = (3,2)
 
-dims = [ 10, 40, 20, 80 ]
-colors = [ "k", "tab:orange", "tab:blue", "tab:green" ]
+colors = [ "k", "tab:blue", "tab:orange", "tab:green" ]
+dims = [ 10, 20, 40, 80 ]
+
+# quantum error scales achievable with the method in newton1968measurability,
+#   optimized over all polar angles
+old_scales = { 10 : 5.205453968693979,
+               20 : 8.689567681451933,
+               40 : 14.26455397815116,
+               80 : 23.36684910939962 }
 
 ##################################################
 
@@ -32,6 +39,9 @@ for dim, color in zip(dims, colors):
     excess_axes = axes - (2*dim-1)
     plt.plot(excess_axes/dim, scales/scales[0],
              ".", label = f"$d={dim}$", color = color)
+
+    old_scale = old_scales[dim] * np.sqrt(2*dim-1)
+    plt.axhline(old_scale/scales[0], color = color, linewidth = 1, zorder = 0)
 
 plt.gca().set_ylim(bottom = 0)
 plt.gca().tick_params(right = True)

### 2021-07-28 sun_phases/scripts/plot_tomo_spheres.py

@@ -5,27 +5,33 @@ import numpy as np
 import scipy, scipy.stats
 import matplotlib.pyplot as plt
 
-from dicke_methods import plot_dicke_state
-from multilevel_methods import drive_op
+from dicke_methods import plot_dicke_state, coherent_spin_state
 
 np.set_printoptions(linewidth = 200)
 
 fig_dpi = 1000
-grid_size = 200
 fig_dir = "../figures/qudit_tomo/"
 
 ##################################################
-# plot a random state
 
-dim = 50
-state_seed = 857054
-np.random.seed(state_seed)
+dim = 10
+kwargs = dict( single_sphere = True, shade = False, grid_size = 501 )
 
-unitary = scipy.stats.unitary_group.rvs(dim)
-state = np.zeros(dim)
-state[-1] = 1
-state = unitary @ state
+# plot a random
+np.random.seed(7)
+random_U = scipy.stats.unitary_group.rvs(dim)
+qudit_state = random_U @ coherent_spin_state("+X", dim-1)
+figure, axes = plot_dicke_state(qudit_state, view_angles = (0,0), **kwargs)
 
-plot_dicke_state(state, grid_size = grid_size)
-plt.savefig(fig_dir + "rnd_state.pdf", dpi = fig_dpi)
-plt.close("all")
+# plot points
+np.random.seed(1)
+point_num = 2*dim-1
+polar = np.arccos(2 * np.random.rand(point_num) - 1)
+azimuth = 2*np.pi*np.random.rand(point_num)
+points = np.array([ np.sin(polar) * np.cos(azimuth),
+                    np.sin(polar) * np.sin(azimuth),
+                    np.cos(polar) ])
+in_front = points[0,:] >= 0
+axes[0].plot(*points[:,in_front], "o", markersize = 15, zorder = 5)
+
+plt.savefig(fig_dir + "sphere_points.pdf", dpi = fig_dpi)

### 2021-07-28 sun_phases/spin_model_paper/main.tex
added minor comment to spin model paper

@@
 \end{enumerate*}
-We construct $g^{(M)}$ for the single-body ($M=1$) case in Appendix \ref{sec:eigenstates}, and provide explicit forms of $g^{(M)}$ with arbitrary $M$.
+We construct $g^{(M)}$ for the single-body ($M=1$) case below (in Appendix \ref{sec:eigenstates}), and provide explicit forms of $g^{(M)}$ with arbitrary $M$.
 
@@
 This Lax analysis correctly predicts that $h_\crit=1$ when $n=2$, but otherwise predicts $h_\crit\sim n^{-3/2}$, which is inconsistent with the finding that $h_\crit\sim n^{-1/3}$ in the mean-field results of the main text (see Figure \ref{fig:crit_fields_X}).
+We emphasize that this inconsistency is not a failure of the Lax formalism, but rather an indication that new theoretical tools are necessary to understand multilevel spin models.
 

### 2021-07-29 sun_phases/spin_qudit_tomo.tex
moved qudit tomography codes to a different repository (https://github.com/perlinm/spin_qudit_tomography) added appendix on optimizing the method of Newton+Young (newton1968measurability) paper, as well as a corresponding discussion in the main text

@@
 The capability to determine upper bounds on reconstruction error a priori motivates a simple randomized tomography protocol that we outline in Section \ref{sec:protocol}, and for which we numerically find that using more measurement axes yields substantial benefits that plateau after $r\approx3d$.
+To facilitate the use of our protocols, we make all of our codes publicly available at Ref.~\cite{tomo_codes}, which also contains the best measurement axes we found for $d\le20$ and $r=3d$.
 
@@
 \end{enumerate}
-Th expected reconstruction error in $\tilde\rho_V$, or its root-mean-square distance from $\rho$, is provided by Eq.~\eqref{eq:error_cls}.
+The expected reconstruction error in $\tilde\rho_V$, or its root-mean-square distance from $\rho$, is provided by Eq.~\eqref{eq:error_cls}.
 If measurement error is shot-noise-limited, then the error in $\tilde\rho_V$ is approximately $\E_V\p{\rho}\approx\E_V\p{\tilde\rho_V}$ and can be computed from Eq.~\eqref{eq:error_qnt}.
@@
     The rapid initial drop in $\tilde\beta(p)$ implies that using more measurement axes can substantially lower the upper bound on reconstruction error provided in Eq.~\eqref{eq:bound}, and that these benefits plateau after $p\approx d$.
-    Horizontal reference lines mark the corresponding (normalized) minimal quantum error scales $\min_\theta\epsilon_{V(\theta)}/\tilde\beta(0)$ achievable with the method in Ref.~\cite{newton1968measurability}, which measures along $2d-1$ axes $V(\theta)$ equally spaced at a polar angle $\theta$.}
+    Horizontal reference lines mark the smallest measurement-adjusted error scales $\min_\theta\beta_\theta/\tilde\beta(0)$ achievable with the method in Ref.~\cite{newton1968measurability}, which is parameterized by an angle $\theta$.}
   \label{fig:axes}
@@
 In the interest of reducing experimental complexity as well as the runtime of our randomized tomography protocol, which grow linearly in $\abs{V}$, we therefore conclude that this protocol should be performed with $\abs{V}\approx3d$ measurement axes.
-
-% todo: paragraph on old method's appearance in fig:axes
-
-% todo: appendix on optimizing the old method that appears in fig:axes
-
-% todo: include table of good measurement axes for specific values of d
-% Yb-173 (6), Cr (7), Sr-87 (10), Er-168 (13), Dy (17), Er-167 (20)
+We provide the best measurement axes that we found for a randomized tomography protocol with $d\le20$ and $\abs{V}=3d$ in Ref.~\cite{tomo_codes}.
+
+For reference, Figure \ref{fig:axes} also shows the smallest measurement-adjusted error scales $\beta_\theta$ achievable with the method of Ref.~\cite{newton1968measurability}, which is comparable to those achieved with our randomized protocol at $\abs{V}\approx3d$ ($p\approx d$).
+The method of Ref.~\cite{newton1968measurability} requires choosing an angle $\theta$, and provides no prescription for making this choice.
+We therefore find the optimal choice of $\theta$ by minimizing $\beta_\theta$ over all $\theta$ (see Appendix \ref{sec:angle_search}), and show $\min\beta_\theta/\tilde\beta(0)$ in Figure \ref{fig:axes}.
+Empirically, we find that the optimal angle for the method of Ref.~\cite{newton1968measurability} is $\theta_{\t{opt}}\approx\frac{\pi}{2}(1 - \frac1{1.34d})$ (see Appendix \ref{sec:angle_search}), although we note that a randomized tomography protocol is likely more robust to errors in axis orientation because measurement axes are generally more spaced out.
+We leave the question of robustness to errors in axis orientation and the effect of these errors on state reconstruction error to future work.
 
@@
 We leave this minimization problem to future work, and note that solving it will likely require making use of the positivity conditions derived in Ref.~\cite{kryszewski2006positivity}.
-A loose lower bound on $\norm{\rho-\sigma_V^\star}$ can be found by minimization under the constraint $\norm{\rho}\le1$, which may provide a tighter upper bound on $\E_V\p{\rho}$ than that in Eq.~\eqref{eq:bound} of the main text.
+A loose lower bound on $\norm{\rho-\sigma_V^\star}_V$ can be found by minimization under the constraint $\norm{\rho}\le1$, which may provide a tighter upper bound on $\E_V\p{\rho}$ than that in Eq.~\eqref{eq:bound} of the main text.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Optimizing the method of Newton and Young}
+\label{sec:angle_search}
+
+Ref.~\cite{newton1968measurability} constructs an explicit protocol for spin qudit tomography, which involves measuring spin projection along $2d-1$ axes equally spaced at a polar angle $\theta$.
+However, this method does not provide any prescription for choosing $\theta$.
+Here, we show the importance of making a good choice of $\theta$, and empirically find the optimal value of $\theta_{\t{opt}}$ that minimizes the corresponding quantum error scale $\epsilon_\theta$, which controls state reconstruction error.
+To this end, Figure \ref{fig:angle_sweep} shows the quantum error scale $\epsilon_\theta$ as a function of the polar angle $\theta$ in the tomography method of Ref.~\cite{newton1968measurability} for a few qudit dimensions $d$.
+While a good choice of $\theta$ yields an error scale $\epsilon_\theta\approx d$ (for the dimensions shown), this error scale can increase by orders of magnitude for poor choices of $\theta$.
+In turn, Figure \ref{fig:opt_angles} shows the optimal angle $\theta_{\t{opt}}$ as a function of the qudit dimension $d$, together with a fit to $\theta_{\t{opt}}=\frac{\pi}{2}(1-\frac1{xd})$ finding $x\approx1.34\pm0.008$.
+
+\begin{figure}
+  \centering
+  \includegraphics{angle_sweep.pdf}
+  \caption{Quantum error scale $\epsilon_\theta$ as a function of the polar angle $\theta$ in the tomography method of Ref.~\cite{newton1968measurability} for a few qudit dimensions $d$.}
+  \label{fig:angle_sweep}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics{opt_angles.pdf}
+  \caption{Optimum angle $\theta_{\t{opt}}$ as a function of qudit dimension $d$ for the tomography method of Ref.~\cite{newton1968measurability}, and a fit to $\theta_{\t{opt}}=\frac{\pi}{2}(1-\frac1{xd})$ finding $x\approx1.34\pm0.008$.}
+  \label{fig:opt_angles}
+\end{figure}
+
 \end{document}

### 2021-07-29 sun_phases/scripts/compute_tomo_times.py

@@ -1,120 +1,0 @@
-#!/usr/bin/env python3
-
-import os, sys, time
-import numpy as np
-
-genesis = time.time()
-
-# flags for what to compute:
-CB = "CB" # classical error bound
-QB = "QB" # quantum error bound
-RE = "RE" # reconstruction error
-
-compute = RE # default
-for idx, arg in enumerate(sys.argv):
-    if arg in [ CB, QB, RE ]:
-        compute = arg
-        sys.argv.remove(arg)
-
-save_data = "save" in sys.argv
-if save_data: sys.argv.remove("save")
-
-# get minimum / maximum qudit dimensions from command-line inputs
-min_dim = int(sys.argv[1])
-try: max_dim = int(sys.argv[2])
-except: max_dim = min_dim
-
-sample_cap = 1000 # maximum number of times we choose a random set of measurement axes
-time_cap = 300 # maximum time to run per qubit dimension, in seconds
-
-seed = 1 # set random number seed. change at your own peril
-np.random.seed(seed)
-
-# determine the directory and file for saving data
-data_dir = "../data/qudit_tomo/"
-if not os.path.isdir(data_dir):
-    os.makedirs(data_dir)
-data_file = data_dir + f"times_{compute}_d{min_dim}-{max_dim}.txt"
-
-while os.path.isfile(data_file):
-    print("file exists:", data_file)
-    data_file += ".new"
-
-##########################################################################################
-# import qudit tomography methods
-
-import qudit_tomo_methods as tomo
-
-random_axes = tomo.random_axes
-meas_mat, meas_scales = tomo.get_meas_methods(max_dim)
-
-if compute == CB:
-    error_scale = tomo.classical_error_scale
-if compute == QB:
-    error_scale = tomo.quantum_error_scale
-if compute == RE:
-    inv_struct_bands = tomo.get_struct_method(max_dim)
-    random_state = tomo.random_state
-    recon_error = tomo.recon_error
-
-print("precomputation finished")
-
-##########################################################################################
-# simulate!
-
-if save_data:
-    # initialize data file
-    with open(data_file, "w") as file:
-        file.write(f"# sample_cap: {sample_cap}\n")
-        file.write(f"# time_cap: {time_cap} sec\n")
-        file.write(f"# seed: {seed}\n")
-        file.write(f"# dim, mean_time")
-        if compute in [ CB, QB ]:
-            file.write(f", error_scale")
-        file.write(f"\n")
-
-for dim in range(min_dim, max_dim+1):
-
[…]

### 2021-07-29 sun_phases/scripts/plot_tomo_axes.py

@@ -1,54 +1,0 @@
-#!/usr/bin/env python3
-
-import os, glob
-import numpy as np
-import matplotlib.pyplot as plt
-
-data_dir = "../data/qudit_tomo/"
-fig_dir = "../figures/qudit_tomo/"
-
-if not os.path.isdir(fig_dir):
-    os.makedirs(fig_dir)
-
-params = { "font.size" : 10,
-           "text.usetex" : True }
-plt.rcParams.update(params)
-
-figsize = (3,2)
-
-colors = [ "k", "tab:blue", "tab:orange", "tab:green" ]
-dims = [ 10, 20, 40, 80 ]
-
-# quantum error scales achievable with the method in newton1968measurability,
-#   optimized over all polar angles
-old_scales = { 10 : 5.205453968693979,
-               20 : 8.689567681451933,
-               40 : 14.26455397815116,
-               80 : 23.36684910939962 }
-
-##################################################
-
-plt.figure(figsize = figsize)
-
-for dim, color in zip(dims, colors):
-    files = glob.glob(data_dir + f"axes_d{dim}.txt")
-    assert(len(files) == 1)
-    axes, _, scales = np.loadtxt(files[0], unpack = True)
-
-    scales *= np.sqrt(axes)
-    excess_axes = axes - (2*dim-1)
-    plt.plot(excess_axes/dim, scales/scales[0],
-             ".", label = f"$d={dim}$", color = color)
-
-    old_scale = old_scales[dim] * np.sqrt(2*dim-1)
-    plt.axhline(old_scale/scales[0], color = color, linewidth = 1, zorder = 0)
-
-plt.gca().set_ylim(bottom = 0)
-plt.gca().tick_params(right = True)
-
-plt.xlabel(r"$p/d$")
-plt.ylabel(r"$\tilde\beta(p)/\tilde\beta(0)$")
-spacing_kwargs = dict( handlelength = 1, columnspacing = 1, labelspacing = 0.2 )
-plt.legend(loc = "best", ncol = 2, **spacing_kwargs)
-plt.tight_layout()
-plt.savefig(fig_dir + "qudit_axes.pdf")

### 2021-07-29 sun_phases/scripts/plot_tomo_times.py

@@ -1,53 +1,0 @@
-#!/usr/bin/env python3
-
-import os, glob
-import numpy as np
-import matplotlib.pyplot as plt
-
-data_dir = "../data/qudit_tomo/"
-fig_dir = "../figures/qudit_tomo/"
-
-if not os.path.isdir(fig_dir):
-    os.makedirs(fig_dir)
-
-params = { "font.size" : 10,
-           "text.usetex" : True }
-plt.rcParams.update(params)
-
-figsize = (3,2)
-size_fits = 20
-
-##################################################
-
-def max_dim(file):
-    pieces = file.replace(".txt","").split("/")[-1].split("_")
-    return pieces[-1].split("-")[-1]
-def get_data(tag):
-    files = sorted(glob.glob(data_dir + f"times_{tag}_d*.txt"), key = max_dim)
-    data = np.vstack([ np.loadtxt(file) for file in files ])
-    return data[:,0], data[:,1]
-
-plt.figure(figsize = figsize)
-
-plot_params = [ ( "CB", r"$\mathcal{S}_V$", "o", "k" ),
-                ( "QB", r"$\epsilon_V$", ".", "tab:blue" ),
-                ( "RE", r"$\mathcal{E}_V(\rho)$", ".", "tab:orange" ) ]
-
-for tag, label, marker, color in plot_params:
-    sizes, times = get_data(tag)
-    plt.loglog(sizes, times, marker, color = color, label = label)
-
-    size_lims = plt.gca().get_xlim()
-    time_lims = plt.gca().get_ylim()
-    fit_idx = slice(-size_fits,-1)
-    fit_args = [ np.log(sizes[fit_idx]), np.log(times[fit_idx]) ]
-    fit, cov = np.polyfit(*fit_args, deg = 1, cov = True)
-    plt.loglog(size_lims, np.exp(fit[0]*np.log(size_lims) + fit[1]), "--", color = color)
-    plt.gca().set_xlim(size_lims)
-    plt.gca().set_ylim(time_lims)
-
-plt.xlabel(r"qudit dimension $(d)$")
-plt.ylabel(r"seconds $(t)$")
-plt.legend(loc = "best")
-plt.tight_layout()
-plt.savefig(fig_dir + "qudit_times.pdf")

### 2021-07-29 sun_phases/scripts/qudit_tomo_methods.py

@@ -1,247 +1,0 @@
-#!/usr/bin/env python3
-
-import sys, time, functools
-import scipy, scipy.linalg
-import numpy as np
-
-import wigner, py3nj
-
-_update = True
-
-##########################################################################################
-# general simulation methods
-
-# get squared singular values of a matrix
-def get_svdvals(matrix):
-    # return scipy.linalg.svdvals(matrix) # for some reason this is slower...
-    if matrix.shape[0] < matrix.shape[1]:
-        MM = matrix @ matrix.conj().T
-    else:
-        MM = matrix.conj().T @ matrix
-    eigvals = scipy.linalg.eigvalsh(MM)
-    eigvals[np.isclose(eigvals,0)] = 0
-    return np.sqrt(eigvals)
-
-# get diagonal bands of a matrix
-def diagonals(mat, band_min = None, band_max = None):
-    rows, cols = mat.shape
-    fill = np.zeros(((cols - 1), cols), dtype = mat.dtype)
-    stacked = np.vstack((mat, fill, mat))
-    major_stride, minor_stride = stacked.strides
-    strides = major_stride, minor_stride * (cols + 1)
-    shape = (rows + cols - 1, cols)
-    reversed_diags = np.lib.stride_tricks.as_strided(stacked, shape, strides)
-    diags = np.roll(np.flipud(reversed_diags), 1, axis = 0)
-    if band_min == None: band_min = -cols
-    if band_max == None: band_max = +rows
-    bands = band_max - band_min
-    return np.roll(diags, -band_min, axis = 0)[:bands+1,:]
-
-# generate random axes on the sphere by uniform sampling
-def random_axes(axis_num):
-    axes = np.random.rand(axis_num,2)
-    axes[:,0] = np.arccos(2*axes[:,0]-1) # polar angles
-    axes[:,1] *= 2*np.pi # azimuthal angles
-    return axes
-
-# generate a random (probably unphysical) qudit state,
-# represented by a point within a `dim^2`-dimensional hypersphere
-def random_state(dim):
-    point = np.array([ np.random.normal() for _ in range(dim**2) ])
-    point /= np.sqrt( np.random.exponential() + sum(point**2) )
-    # organize components into (2L+1)-sized vectors of "degree" L < dim
-    return { LL : point[ LL**2 : LL**2 + 2*LL+1 ] for LL in range(dim) }
-
[…]

### 2021-07-29 sun_phases/spin_qudit_tomo.tex
minor updates to tomography paper

@@
 
-\title{Spin qudit tomography}
+\title{Spin qudit tomography and reconstruction error}
 \author{Michael A.~Perlin}
@@
 and offer unique capabilities for quantum error correction \cite{albert2020robust, gross2021designing}.
-In all cases, quantum state tomography is necessary to take full advantage of a spin qudit.
-
-% todo: explain the role of tomography in sensing and simulation?
+In all cases, quantum state tomography is necessary to take full advantage of a spin qudit\footnote{Note that the measurement of collective observables for quantum sensing or simulation can be recast as a single-spin tomography task.}.
 
@@
 The protocols based on infinite-dimensional representations of a quantum spin have the advantage of reconstructing its state from measurements of spin projection onto different spatial axes, which are generally accessible with any spin qudit.
-Nonetheless, in addition to their inefficiencies these protocols obfuscate the minimal requirements for performing full state tomography, provide no straightforward error bounds or guarantees of accuracy, and (with the notable exception of Ref.~\cite{schmied2011tomographic}) generally extract only a small fraction of the information contained in measurement data.
-
-In this work, we consider the task of performing spin qubit tomography using only measurements of spin projection onto different spatial axes, and address some of the shortcomings discussed above.
-This sort of qudit tomography was first considered in Ref.~\cite{newton1968measurability}, as well as a few later works \cite{hofmann2004quantumstate, filippov2010inverse, schmied2011tomographic}.
-Specifically, Ref.~\cite{newton1968measurability} provided an explicit protocol for reconstructing a $d$-level spin qudit state from measurements of spin projection along $2d-1$ axes, the minimum number necessary for full tomography of an arbitrary qudit state.
+Nonetheless, these protocols obfuscate the minimal requirements for performing full state tomography, provide no straightforward error bounds or guarantees of accuracy, and (with the notable exception of Ref.~\cite{schmied2011tomographic}) generally extract only a small fraction of the information contained in measurement data.
+
+In this work, we consider the task of performing spin qubit tomography using only measurements of spin projection onto different spatial axes.
+This sort of task was first considered in Ref.~\cite{newton1968measurability}, as well as a few later works \cite{hofmann2004quantumstate, filippov2010inverse, schmied2011tomographic}.
+Specifically, Ref.~\cite{newton1968measurability} provided an explicit protocol for reconstructing a $d$-level spin qudit state from measurements of spin projection along $2d-1$ axes, the minimum number necessary for full tomography of an arbitrary (possibly mixed) qudit state.
 However, the protocol in Ref.~\cite{newton1968measurability} involves a choice of a single (arbitrary) angle $\theta$, and provides no means for comparing different choices of $\theta$, which may result in wildly different statistical errors (i.e.~precision) in a reconstructed state.
@@
 \begin{enumerate*}
-\item require making assumptions about the qudit state in question \cite{schmied2011tomographic},
+\item require making assumptions about the qudit state in question \cite{schmied2011tomographic} (making the tomographic protocol only valid for a restricted set of possible states),
 \item do not address the question of statistical error \cite{hofmann2004quantumstate}, or
-\item provide a measure of statistical error that is overly conservative and computationally demanding \cite{filippov2010inverse}.
+\item provide a measure of statistical error that is needlessly conservative and computationally demanding \cite{filippov2010inverse}.
 \end{enumerate*}
+We address these shortcomings in this work, and identify remaining avenues for refining spin qudit tomography protocols.
 
@@
 The capability to determine upper bounds on reconstruction error a priori motivates a simple randomized tomography protocol that we outline in Section \ref{sec:protocol}, and for which we numerically find that using more measurement axes yields substantial benefits that plateau after $r\approx3d$.
-To facilitate the use of our protocols, we make all of our codes publicly available at Ref.~\cite{tomo_codes}, which also contains the best measurement axes we found for $d\le20$ and $r=3d$.
+To facilitate the use of our protocols, we make all of our codes publicly available at Ref.~\cite{tomo_codes}, which also contains the best measurement axes we found for $d\le30$ and $r=3d$.
 
@@
 \end{align}
-Here $\bm v$ and $\ell$ label a row of $M_V$, or equivalently label a standard basis vector $\ket{\bm v\ell}$ of a $\p{\abs{V}\times d}$-dimensional vector space, and $\obra{T_{\bm v\ell,0}}$ is the conjugate transpose of $\oket{T_{\bm v\ell,0}}$.
-A necessary and sufficient condition for $V$ to allow for full state tomography is that the measured polarization operators in $\set{T_{\bm v\ell,0}}$, or equivalently the columns of $M_V$, span the entire ($d^2$-dimensional) space of operators on a $d$-level spin qudit.
+Here $\bm v$ and $\ell$ label a row of $M_V$, or equivalently label a standard (``one-hot'') basis vector $\ket{\bm v\ell}$ of a $\p{\abs{V}\times d}$-dimensional vector space, and $\obra{T_{\bm v\ell,0}}$ is the conjugate transpose of $\oket{T_{\bm v\ell,0}}$.
+A necessary and sufficient condition for $V$ to allow for full state tomography is that the measured polarization operators $T_{\bm v\ell,0}$, or equivalently the columns of $M_V$, span the entire ($d^2$-dimensional) space of operators on a $d$-level spin qudit.
 In this case $M_V$ must be full rank, with $d^2$ nonzero singular values.
@@
 \end{align}
-where, using the fact that $T_j=T_j^\dag$ (because they are degree $m=0$ polarization operators),
+where, using the fact that $T_j=T_j^\dag$ (because they are diagonal, degree $m=0$ polarization operators),
 \begin{align}
@@
 The measurement matrix $M_V$ allows us to make concrete statements about the statistical error between the empirical estimate $\tilde\rho_V$ and the true state $\rho$.
-Assume, for example, that the estimates $\tilde T_j$ are equal to $\bk{T_j}_\rho$ up to uncorrelated noise with variance at most $\epsilon^2$:
+Assume, for example, that the estimates $\tilde T_j$ are equal to $\bk{T_j}_\rho$ up to uncorrelated noise with variance no grater than $\epsilon^2$:
 \begin{align}
@@
   &&
-  \bbk{\epsilon_j \epsilon_{j'}} \le \epsilon^2 \delta_{jj'},
+  \bbk{\epsilon_j \epsilon_{j'}} \le \epsilon^2 \delta_{jj'}.
   \label{eq:error_assumption}
 \end{align}
-where $\set{\epsilon_j}$ are independent random variables and we use the double brackets $\bbk{\cdot}$ to denote statistical averaging over experimental trials that estimate $\bk{T_j}_\rho$.
+Here $\set{\epsilon_j}$ are independent random variables, and we use the double brackets $\bbk{\cdot}$ to denote statistical averaging over experimental trials that estimate $\bk{T_j}_\rho$.
 In this case, the mean squared error with which $\tilde\rho^V_k$ approximates $\rho^V_k$ is
@@
 \end{align}
-where $\norm{X}^2 \equiv \obk{X|X} = \tr\p{X^\dag X}$ is the squared (Frobenius, or Hilbert-Schmidt) norm of $X$, and the {\it classical error scale} $\S_V$ is defined by
+where $\norm{X}^2 \equiv \obk{X|X} = \tr\p{X^\dag X}$ is the squared (Euclidean, Frobenius, or Hilbert-Schmidt) norm of $X$, and the {\it classical error scale} $\S_V$ is defined by
 \begin{align}
@@
 If measurement error is shot-noise-limited, then the error in $\tilde\rho_V$ is approximately $\E_V\p{\rho}\approx\E_V\p{\tilde\rho_V}$ and can be computed from Eq.~\eqref{eq:error_qnt}.
-If $\tilde\rho_V$ has negative eigenvalues, its distance from $\rho$ can be reduced with maximum-likelihood corrections \cite{smolin2012efficient}.
-
-In principle, any information about $\rho$ obtained from prior knowledge or preliminary measurement data can be used to construct tailored or adaptive measurement protocols \cite{huszar2012adaptive, ferrie2014selfguided, granade2016practical, pereira2018adaptive} that are more efficient in terms of the number of measurements required to estimate $\rho$ to a fixed precision.
-We leave the development of tailored and adaptive measurement protocols to future work.
+If $\tilde\rho_V$ has negative eigenvalues, its distance from $\rho$ can be reduced with maximum-likelihood corrections \cite{smolin2012efficient}, which will additionally guarantee that $\tilde\rho_V$ satisfies all requirements for being a physical state.
 
@@
     The rapid initial drop in $\tilde\beta(p)$ implies that using more measurement axes can substantially lower the upper bound on reconstruction error provided in Eq.~\eqref{eq:bound}, and that these benefits plateau after $p\approx d$.
-    Horizontal reference lines mark the smallest measurement-adjusted error scales $\min_\theta\beta_\theta/\tilde\beta(0)$ achievable with the method in Ref.~\cite{newton1968measurability}, which is parameterized by an angle $\theta$.}
+    Horizontal reference lines mark the smallest measurement-adjusted error scales $\min_\theta\beta_\theta/\tilde\beta(0)$ achievable with the method in Ref.~\cite{newton1968measurability}, which is parameterized by an arbitrarily chosen angle $\theta$.}
   \label{fig:axes}
@@
   \beta(p)^2
-  \equiv \min_V\Set{\epsilon_V^2\abs{V}:\abs{V}=2d-1+p}
-\end{align}
-as a function of $p$, the number of measurement axes exceeding $2d-1$.
+  \equiv \min_V\Set{\epsilon_V^2\abs{V}:\abs{V}=2d-1+p},
+\end{align}
+where $p$ is the number of ``extra'' measurement axes exceeding $2d-1$.
 Though we cannot minimize over all suitable choices of measurement axes $V$ to compute $\beta(p)$, we can compute an empirical upper bound $\tilde\beta(p)\ge\beta(p)$ by minimizing over a large number of randomly chosen $V$.
@@
 In the interest of reducing experimental complexity as well as the runtime of our randomized tomography protocol, which grow linearly in $\abs{V}$, we therefore conclude that this protocol should be performed with $\abs{V}\approx3d$ measurement axes.
-We provide the best measurement axes that we found for a randomized tomography protocol with $d\le20$ and $\abs{V}=3d$ in Ref.~\cite{tomo_codes}.
-
-For reference, Figure \ref{fig:axes} also shows the smallest measurement-adjusted error scales $\beta_\theta$ achievable with the method of Ref.~\cite{newton1968measurability}, which is comparable to those achieved with our randomized protocol at $\abs{V}\approx3d$ ($p\approx d$).
+We provide the best measurement axes that we found for a randomized tomography protocol with $d\le30$ and $\abs{V}=3d$ in Ref.~\cite{tomo_codes}.
+
+For reference, Figure \ref{fig:axes} also shows the smallest measurement-adjusted error scales $\beta_\theta$ achievable with the method of Ref.~\cite{newton1968measurability}, which is comparable to those achieved with our randomized protocol at $\abs{V}\approx3d$.
 The method of Ref.~\cite{newton1968measurability} requires choosing an angle $\theta$, and provides no prescription for making this choice.
-We therefore find the optimal choice of $\theta$ by minimizing $\beta_\theta$ over all $\theta$ (see Appendix \ref{sec:angle_search}), and show $\min\beta_\theta/\tilde\beta(0)$ in Figure \ref{fig:axes}.
+We therefore find the optimal choice of $\theta$ by minimizing the error scale $\beta_\theta$ over all $\theta$ (see Appendix \ref{sec:angle_search}), and show $\min\beta_\theta/\tilde\beta(0)$ in Figure \ref{fig:axes}.
 Empirically, we find that the optimal angle for the method of Ref.~\cite{newton1968measurability} is $\theta_{\t{opt}}\approx\frac{\pi}{2}(1 - \frac1{1.34d})$ (see Appendix \ref{sec:angle_search}), although we note that a randomized tomography protocol is likely more robust to errors in axis orientation because measurement axes are generally more spaced out.
 We leave the question of robustness to errors in axis orientation and the effect of these errors on state reconstruction error to future work.
+
+As a final point, we note that any information about an unknown qudit state $\rho$, obtained from prior knowledge or preliminary measurement data, can be used to construct tailored or adaptive measurement protocols \cite{huszar2012adaptive, ferrie2014selfguided, granade2016practical, pereira2018adaptive} that are more efficient in terms of the number of measurements required to estimate $\rho$ to a fixed precision.
+We leave the development of tailored and adaptive measurement protocols to future work.
 
@@
 Note that the bound in Eq.~\eqref{eq:bound_apndx} is not tight, as the individual bounds on the variances $\bbk{\epsilon_{i\ell}^2}$ cannot all be achieved simultaneously.
-There is therefore still room for improvement on the bound in Eq.~\eqref{eq:bound} by maximizing $\E_V$ over the set of all qudit states $\rho$.
+There is therefore still room for improvement on the bound in Eq.~\eqref{eq:bound} by maximizing $\E_V$ over the set of physically achievable qudit states $\rho$.
 
@@
 While a good choice of $\theta$ yields an error scale $\epsilon_\theta\approx d$ (for the dimensions shown), this error scale can increase by orders of magnitude for poor choices of $\theta$.
-In turn, Figure \ref{fig:opt_angles} shows the optimal angle $\theta_{\t{opt}}$ as a function of the qudit dimension $d$, together with a fit to $\theta_{\t{opt}}=\frac{\pi}{2}(1-\frac1{xd})$ finding $x\approx1.34\pm0.008$.
+In turn, Figure \ref{fig:opt_angles} shows the optimal angle $\theta_{\t{opt}}$ as a function of the qudit dimension $d$, together with a fit to $\theta_{\t{opt}}=\frac{\pi}{2}(1-\frac1{xd})$ finding $x\approx1.34$.
 
@@
   \includegraphics{opt_angles.pdf}
-  \caption{Optimum angle $\theta_{\t{opt}}$ as a function of qudit dimension $d$ for the tomography method of Ref.~\cite{newton1968measurability}, and a fit to $\theta_{\t{opt}}=\frac{\pi}{2}(1-\frac1{xd})$ finding $x\approx1.34\pm0.008$.}
+  \caption{Optimum angle $\theta_{\t{opt}}$ as a function of qudit dimension $d$ for the tomography method of Ref.~\cite{newton1968measurability}, and a fit to $\theta_{\t{opt}}=\frac{\pi}{2}(1-\frac1{xd})$ finding $x\approx1.34$.}
   \label{fig:opt_angles}

### 2021-08-30 sun_phases/spin_qudit_tomo.tex
minor changes to tomography paper

@@
 
-\title{Spin qudit tomography and reconstruction error}
+\title{Spin qudit tomography and state reconstruction error}
 \author{Michael A.~Perlin}
@@
 
-Due to advancements in experimental capabilities to address nuclear spin states (i.e.~hyperfine levels) in ultracold atomic systems \cite{daley2011quantum, mischuck2012control, smith2013quantum, cazalilla2014ultracold}, as well as developments in the control of ultracold molecular systems \cite{bohn2017cold, zeppenfeld2012sisyphus, kozyryev2017sisyphus, puri2017synthesis, wu2017cryofuge, marco2019degenerate, liu2019molecular, anderegg2019optical, chou2020frequencycomb, lin2020quantum}, a particular setting of growing interest is the spin qudit, or a multilevel quantum angular momentum degree of freedom.
+Due to advancements in experimental capabilities to address nuclear spin states (i.e.~hyperfine levels) in ultracold atomic systems \cite{daley2011quantum, lu2011strongly, mischuck2012control, aikawa2012boseeinstein, smith2013quantum, cazalilla2014ultracold}, as well as developments in the control of ultracold molecular systems \cite{bohn2017cold, takekoshi2014ultracold, zeppenfeld2012sisyphus, kozyryev2017sisyphus, puri2017synthesis, wu2017cryofuge, marco2019degenerate, liu2019molecular, anderegg2019optical, chou2020frequencycomb, lin2020quantum}, a particular setting of growing interest is the spin qudit, or a multilevel quantum angular momentum degree of freedom.
 Spin qudits can provide advantages over their qubit counterparts for quantum sensing \cite{hemmer2018squeezing, evrard2019enhanced}, enable quantum simulations of SU($d$) magnetism \cite{cazalilla2014ultracold, banerjee2013atomic, zhang2014spectroscopic, scazza2014observation, goban2018emergence}, % todo: cite SU(n) spin model paper!
@@
 
-The problem of qudit tomography is not new, with an extensive literature on a variety of techniques \cite{newton1968measurability, hofmann2004quantumstate, filippov2010inverse, schmied2011tomographic, evrard2019enhanced, flammia2005minimal, thew2002qudit, salazar2012quantum, sosa-martinez2017quantum, ha2018minimal, stefano2019set, palici2020oam, filippov2010inverse, hofmann2004quantumstate, filippov2010inverse}.
+The problem of qudit tomography is not new, with an extensive literature on a variety of techniques \cite{newton1968measurability, hofmann2004quantumstate, filippov2010inverse, schmied2011tomographic, evrard2019enhanced, flammia2005minimal, thew2002qudit, salazar2012quantum, sosa-martinez2017quantum, ha2018minimal, stefano2019set, palici2020oam}.
 However, most existing protocols either rely on infinite-dimensional representations of a quantum spin \cite{manko1997spin, schmied2011tomographic, evrard2019enhanced}, or require the capability to perform essentially arbitrary operations on a qudit \cite{thew2002qudit, flammia2005minimal, salazar2012quantum, sosa-martinez2017quantum, ha2018minimal, stefano2019set, palici2020oam}, generally resulting in tomographic protocols that can be highly inefficient or unachievable in practice.
@@
 \end{align}
-where $\bk{s\mu;\ell m|s\nu}$ is a Clebsh-Gordan coefficient that enforces $\ell\in\set{0,1,\cdots,d-1}$ and $m\in\set{-\ell,-\ell+1,\cdots,\ell}$, such that there are $d^2$ polarization operators in total.
+where $\ket\mu$ is an eigenstate of the axial spin projection operator $S_\z\ket\mu=\mu\ket\mu$; and $\bk{s\mu;\ell m|s\nu}$ is a Clebsh-Gordan coefficient that enforces $\ell\in\set{0,1,\cdots,d-1}$ and $m\in\set{-\ell,-\ell+1,\cdots,\ell}$, such that there are $d^2$ polarization operators in total.
 For brevity, we will generally treat the value of $d$ as constant but arbitrary throughout this work, and we will suppress any explicit dependence of quantities or operators such as $T_{\ell m}$ on $d$.
@@
 We are thus nominally restricted to measuring projectors $\Pi_{\bm v\mu} \equiv \op{\mu_{\bm v}}$, where $\ket{\mu_{\bm v}}\equiv R\p{\bm v}\ket{\mu}$ is a state with spin projection $\mu$ onto the measurement axis $\bm v$.
-For any fixed axis $\bm v$, the sets $\set{\Pi_{\bm v\mu}}$ and $\set{T_{\bm v\ell,0}}$ are both complete bases for the space of operators that are diagonal in the basis $\set{\ket{\mu_{\bm v}}}$.
+For any fixed axis $\bm v$, the sets $\set{\Pi_{\bm v\mu}}$ and $\set{T_{\bm v\ell,0}}$ (i.e.~all $T_{\bm v\ell m}$ with $m=0$) are both complete bases for the space of operators that are diagonal in the basis $\set{\ket{\mu_{\bm v}}}$.
 Measuring the projectors $\set{\Pi_{\bm v\mu}}$ is therefore equivalent to measuring the polarization operators $\set{T_{\bm v\ell,0}}$, and provides data on the expectation values $\bk{T_{\bm v\ell,0}}_\rho$.
 
-In order to reconstruct an arbitrary density operator $\rho$ from the expectation values $\bk{T_{\bm v\ell,0}}_\rho$, we essentially need to to find a set of coefficients $C_{\ell mk}\p{\bm v}$ that would allow us to recover any matrix element $\rho_{\ell m}$ of $\rho$ through
+In order to reconstruct an arbitrary density operator $\rho$ from the expectation values $\bk{T_{\bm v\ell,0}}_\rho$, we essentially need to find a set of coefficients $C_{\ell mk}\p{\bm v}$ that would allow us to recover any matrix element $\rho_{\ell m}$ of $\rho$ through
 \begin{align}
@@
 
-% todo: look up and cite randomized protocols
-% see JQI / Jake Taylor / Mohammad Hafezi
-
 The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following protocol for spin qudit tomography:
@@
   \bbk{\epsilon_\O \epsilon_\Q}
-  = \bbk{\p{\tilde\O - \bk{\O}} \p{\tilde\Q - \bk{\Q}}}
+  = \Bbk{\p{\tilde\O - \bk{\O}} \p{\tilde\Q - \bk{\Q}}}
   = \bk{\p{\O - \bk{\O}} \p{\Q - \bk{\Q}}}
@@
 \end{align}
-If $\p{i,j}\ne\p{i',j'}$, then $\Pi_{i\mu}^j$ and $\Pi_{i'\mu'}^{j'}$ address different tensor factors of the product state $\rho^{\otimes N}$, so the expectation value of their product factorizes due to the identity $\tr\sp{\p{A\otimes B}\p{A'\otimes B'}} = \tr\p{AA'}\times\tr\p{BB'}$.
+If $\p{i,j}\ne\p{i',j'}$, then $\Pi_{i\mu}^j$ and $\Pi_{i'\mu'}^{j'}$ address different tensor factors of the product state $\rho^{\otimes N}$, so the expectation value of their product factorizes due to the fact that $\tr\sp{\p{A\otimes B}\p{A'\otimes B'}} = \tr\p{AA'}\times\tr\p{BB'}$.
 This factorization can also be seen as a consequence of the fact that if $\p{i,j}\ne\p{i',j'}$, then $\Pi_{i\mu}^j$ and $\Pi_{i'\mu'}^{j'}$ are ``spatially separated'' on $\rho^{\otimes N}$, which means that their expectation values cannot have quantum correlations.

### 2021-08-30 sun_phases/spin_model_paper/main.tex
minor tweaks to spin model paper

@@
 \newcommand{\su}{\mathfrak{su}}
-\renewcommand{\ss}{\bar{\v s}\cdot\bar{\v s}}
+\newcommand{\sds}{\bar{\v s}\cdot\bar{\v s}}
 \DeclareMathOperator{\diag}{diag}
@@
 
-\title{Engineering infinite-range SU($n$) interactions with spin-orbit coupled fermions in an optical lattice}
+\title{Engineering infinite-range SU($n$) interactions with spin-orbit-coupled fermions in an optical lattice}
 \author{Michael A. Perlin}
@@
 \author{Diego Barberena}
+\author{Mikhail Mamaev}
+\author{Bhuvanesh Sundar}
 \affiliation{\JILA}
@@
   \H_{\t{kin}}
-  &= -J \sum_{\bk{j,j'},\mu} \c_{j\mu}^\dag \c_{j'\mu} + \t{h.c.}, \\
+  &= -J \sum_{\bk{j,j'},\mu} \c_{j\mu}^\dag \c_{j'\mu} + \t{h.c.},
+  \label{eq:H_kin_FH} \\
   \H_{\t{int}}
@@
   \c_{j\mu}^\dag \c_{j\mu} \c_{j\nu}^\dag \c_{j\nu},
+  \label{eq:H_int_FH}
 \end{align}
@@
 If the interaction energy $U$ is small compared to the single-particle bandwidth $4J$, then the mode-changing collisions in $\H_{\t{int}}$ become off-resonant, motivating the frozen-mode approximation $\set{k,p}=\set{\ell,q}$ (which we assume for the remainder of this work)\footnote{Note that the frozen-mode approximation neglects correlated momentum-hopping terms of the form $\c_{\pi-p,\mu}^\dag \c_{\pi-q,\mu} \c_{p\nu}^\dag \c_{q\nu}$, which conserve both momentum and energy.
-We defer a careful treatment of these terms to future work, noting only that they vanish on the manifold of permutationally symmetric spin states with one atom per lattice site.}.
-The terms with $k=\ell$ and $p=q$ are $\frac{u}{2N} \sum \c_{\ell\mu}^\dag \c_{\ell\mu} \c_{q\nu}^\dag \c_{q\nu} = \frac12 N u$, which is an overall energy shift that we can freely neglect.
+We defer a careful treatment of these terms to future work, noting only that they vanish on the manifold of permutationally symmetric spin states with one atom per lattice site, and that the frozen-mode approximation is benchmarked in Refs.~\cite{he2019engineering, smale2019observation} and Appendix \ref{sec:benchmarking}.}.
+The terms with $k=\ell$ and $p=q$ are $\frac{u}{2N} \sum \c_{\ell\mu}^\dag \c_{\ell\mu} \c_{q\nu}^\dag \c_{q\nu} = \frac12 N u$, which is a constant  energy shift that we can freely neglect.
 Defining the spin operators $\s_{\mu\nu q}\equiv \c_{q\mu}^\dag \c_{q\nu}$, the remaining terms of the kinetic and interaction Hamiltonians are
@@
 In this case we can simply treat our system as $N$ distinguishable $n$-level quantum spins at ``lattice sites'' $p,q$.
-Note that the ``kinetic'' terms of this spin model ($\H_{\t{kin}}$) are proportional to the identity operator, contributing an overall shift in energy that we can neglect and this point.
+Note that the ``kinetic'' terms of this spin model ($\H_{\t{kin}}$) are proportional to the identity operator, contributing an overall shift in energy that we can neglect at this point.
 Nevertheless, these kinetic terms will become important in the presence of an external drive, which we discuss in Section \ref{sec:SOC}.
+The validity of approximating the Hubbard model in Eqs.~\eqref{eq:H_kin_FH}--\eqref{eq:H_int_FH} by the spin model in Eqs.~\eqref{H_kin_start}--\eqref{eq:H_int_start} has been previously benchmarked for SU(2)-symmetric interactions \cite{he2019engineering, smale2019observation}, and we provide additional benchmarking for SU(4) and SU(6) in Appendix \ref{sec:benchmarking}.
 
@@
 
-We assume that all driving lasers are detuned by $\Delta$ below an electronic transition.
-Each laser is indexed by an axis of propagation, $\v v$, and a drive amplitude $\Omega_{\v v,+}$ ($\Omega_{\v v,-}$) for right-circularly (left-circularly) polarized light.
-An electronic $\ket\g\to\ket\e$ excitation is thus accompanied by a nuclear spin transition $\mu_{\v v}\to\mu_{\v v}\pm1$, where $\mu_{\v v}$ is the projection of nuclear spin onto the axis $\v v$.
-Furthermore, a laser pointing along $\v v$ will generally imprint a phase $e^{-\i\kappa\v v\cdot\v\ell j}$ on lattice site $j$, where $\kappa$ is the wavenumber of the laser (in units with lattice spacing $a=1$) and $\v\ell$ is a fixed unit vector parallel to the lattice.
-On a one-dimensional lattice with transverse confinement, the multi-laser drive Hamiltonian is then
-\begin{align}
-  \H_{\t{drive}}^{\t{full}}
-  = \sum_{j,\v v,\sigma} \Omega_{\v v\sigma}
-  \p{e^{-\i\kappa\v v\cdot\v\ell j} \s_{\v v\sigma j}\otimes\op{\e}{\g}_j + \t{h.c.}}
-  + \Delta \hat N_\e,
-  \label{eq:H_drive_full}
-\end{align}
-where $\sigma\in\set{+1,-1}$ indexes a polarization of light; $\s_{\v v\sigma j}$ is an $n$-level spin-raising ($\sigma=+1$) or spin-lowering ($\sigma=-1$) operator along axis $\v v$ for an atom on lattice site $j$ (identical to that for $n-1$ spin-$\frac12$ particles in the Dicke manifold); $\ket\g_j$ and $\ket\e_j$ respectively denote the ground and excited electronic states of atom $j$; and $\hat N_\e=\1\otimes\sum_j\op\e_j$ counts the number of excited atoms (with $\1$ the identity operator on all spin degrees of freedom).
-
 \begin{figure}
@@
 A third, linearly polarized laser with amplitude $\Omega_0$ points in a direction orthogonal to both the lattice and the other driving lasers.
-Absorbing a photon from the laser with amplitude $\Omega_m$ induces a transition $(\g,\mu)\to(\e,\mu+m)$ for the (electronic, nuclear spin) state of an atom, where nuclear spin is quantized along the z axis.
+Absorbing a photon from the laser with amplitude $\Omega_m$ induces a transition $(\g,\mu)\to(\e,\mu+m)$ for the (electronic, nuclear spin) state of an atom, where nuclear spin is quantized along the $z$ axis.
 }
@@
 
-We now consider a specific three-laser driving scheme with a geometry sketched in Figure \ref{fig:3LD}.
-Here the lattice lies in the y-z plane at an angle $\theta$ to the z axis, oriented along $\v\ell=(0,\sin\theta,\cos\theta)$; two counter-propagating right-circularly polarized lasers point in opposite directions along the z axis, $\v v_\pm=(0,0,\pm1)$; and a third linearly-polarized laser points along the x axis, $\v v_0=(1,0,0)$.
-Choosing $\v v_+$ (i.e.~the z axis) as the spin quantization axis, we can define new drive amplitudes $\Omega_0\equiv-(\Omega_{\v v_0,+}+\Omega_{\v v_0,-})$ and $\Omega_\pm\equiv\pm\Omega_{\v v_\pm,+}$, as well as a SOC angle $\phi\equiv\kappa\v v_+\cdot\v\ell=\kappa\cos\theta$, in terms of which the drive Hamiltonian becomes
+We consider a specific three-laser driving scheme with a geometry sketched in Figure \ref{fig:3LD}.
+Here the lattice lies in the $y$-$z$ plane at an angle $\theta$ to the $z$ axis, oriented along $\v\ell=(0,\sin\theta,\cos\theta)$.
+We set the spin quantization axis along $z$.
+The laser setup consists of
+\begin{enumerate*}
+\item two counter-propagating right-circularly polarized lasers with with drive amplitudes $\Omega_\pm$ and wavevectors $\kappa\v v_\pm$, propagating in opposite directions along the $z$ axis, $\v v_\pm=(0,0,\pm1)$, and
+\item a third laser linearly polarized along $z$, with drive amplitude $\Omega_0$ and wavevector $\kappa \v v_0$, propagating along the $x$ axis $\v v_0=(1,0,0)$.
+\end{enumerate*}
+All driving lasers are detuned by $\Delta$ below an electronic transition.
+The full Hamiltonian for this three-laser drive can be written as
 \begin{align}
@@
 \end{align}
-where $m\in\set{+1,0,-1}$ indexes the laser pointing along $\v v_m$; $\s_{0,j}\equiv\s_{\z,j}$ for shorthand; and $\s_{\z,j},\s_{+,j}\s_{-,j}$ are standard axial, spin-raising, and spin-lowering operators for spin $j$.
+where $m\in\set{+1,0,-1}$ indexes the laser pointing along $\v v_m$; the SOC angle $\phi\equiv\kappa\v v_+\cdot\v\ell=\kappa a \cos\theta$ (with $a$ the lattice spacing); $\s_{\z,j},\s_{+,j}\s_{-,j}$ are standard axial, spin-raising, and spin-lowering operators for spin $j$; $\s_{0,j}\equiv\s_{\z,j}$ for shorthand; $\ket\g_j$ and $\ket\e_j$ respectively denote the ground and excited electronic states of atom $j$; and $\hat N_\e=\1\otimes\sum_j\op\e_j$ counts the number of excited atoms (with $\1$ the identity operator on all spin degrees of freedom).
+
 In the far-detuned limit $\abs{\Delta}\gg\abs{\Omega_m}$, a second-order perturbative treatment of electronic excitations ($\ket\e$) yields an effective drive Hamiltonian that only addresses ground-state nuclear spins.
@@
 Initial states are nominally prepared in the ``lab frame'', and must be transformed according to the gauge transformation $\c_{j\mu}^\dag\to e^{\i\phi\mu j} \c_{j\mu}^\dag$ prior to evolution under the three-laser drive $\H_{\t{3LD}}$ in Eq.~\eqref{eq:H_3LD}, which is expressed in the ``gauge frame''.
-We assume the capability to prepare an initial state in which all spins are maximally polarized along the z axis, i.e.~$\ket{\z}^{\otimes N}=\ket{s}^{\otimes N}$, which is unaffected by the gauge transformation (up to a global phase).
+We assume the capability to prepare an initial state in which all spins are maximally polarized along the $z$ axis, i.e.~$\ket{\z}^{\otimes N}=\ket{s}^{\otimes N}$, which is unaffected by the gauge transformation (up to a global phase).
 The three-laser then allows us to rotate this state into one that is polarized along any spatial axis (in the gauge frame).
@@
 In contrast, the drive we considered in Section \ref{sec:controls} addresses electronic excitations off-resonantly, inducing an effective Hamiltonian in the ground-state hyperfine manifold with spin projections $\mu\in\set{s,s-1,\cdots,-s}$ (a similar scheme was used to study SOC in a subspace of the ground-state manifold in Ref.~\cite{mancini2015observation}).
-Nonetheless, both the two-level drive in Eq.~\eqref{eq:drive_2} and the $n$-level drive in Eq.~\eqref{eq:H_3LD} become homogeneous (i.e.~independent of the spatial mode index $j$ or $q$) and independent of the SOC angle $\phi$ after the same spin-symmetric gauge transformation\footnote{Note that the ``asymmetric'' gauge transformation $(\c_{j,\up}^\dag,\c_{j,\dn}^\dag)\to(e^{\i\phi j} \c_{j,\up}^\dag,\c_{j,\dn}^\dag)$, sometimes performed in the two-state SOC literature, does not generalize as nicely to $n>2$.} $\c_{j\mu}^\dag\to e^{\i\phi\mu j} \c_{j\mu}^\dag$.
+Nonetheless, both the two-level drive in Eq.~\eqref{eq:drive_2} and the $n$-level drive in Eq.~\eqref{eq:H_3LD} become homogeneous (i.e.~independent of the spatial mode index $j$ or $q$) and independent of the SOC angle $\phi$ after the same spin-symmetric gauge transformation\footnote{The ``asymmetric'' gauge transformation $(\c_{j,\up}^\dag,\c_{j,\dn}^\dag)\to(e^{\i\phi j} \c_{j,\up}^\dag,\c_{j,\dn}^\dag)$, sometimes performed in the two-state SOC literature, does not generalize as nicely to $n>2$.} $\c_{j\mu}^\dag\to e^{\i\phi\mu j} \c_{j\mu}^\dag$.
 
@@
   \H_{\t{int}}^{(\phi)}
+  \equiv \f{u}{2N} \sum_{k,\ell,p,q,\mu,\nu}
+  \c_{k\mu}^{(\phi)\dag} \c_{\ell\mu}^{(\phi)}
+  \c_{p\nu}^{(\phi)\dag} \c_{q\nu}^{(\phi)}
+  \times \delta_{k+p,\ell+q},
+\end{align}
+and $\c_{k\mu}^{(\phi)}\equiv\c_{k-\mu\phi,\mu}$.
+However, $\H_{\t{int}}^{(\phi)}$ cannot be reduced to a spin Hamiltonian that addresses the same spin degrees of freedom as $\H_{\t{kin}}^{(\phi)}$.
+Again making a frozen-mode approximation, or equivalently keeping only terms that preserve the coherences of an initially spin-polarized state, the gauge-transformed interaction Hamiltonian becomes
+\begin{align}
+  \H_{\t{int}}^{(\phi)}
   \equiv -\f{u}{2N} \sum_{p,q,\mu,\nu}
-  \c_{p-\mu\phi,\mu}^\dag \c_{p-\nu\phi,\nu}
-  \c_{q-\nu\phi,\nu}^\dag \c_{q-\mu\phi,\mu}.
-\end{align}
-However, $\H_{\t{int}}^{(\phi)}$ cannot be exactly reduced to a spin Hamiltonian that addresses the same spin degrees of freedom as $\H_{\t{kin}}^{(\phi)}$.
-Nonetheless, for sufficiently weak SOC (small $s\phi$) the effect of $\H_{\t{int}}^{(\phi)}$ may be approximated by $\H_{\t{int}}^{(0)}=\H_{\t{int}}$, which has previously been benchmarked for SU($2$)-symmetric interactions \cite{he2019engineering, smale2019observation}.
-To ensure that $\H_{\t{kin}}^{(\phi)}$ does not become trivial as $\phi\to0$, we can keep $J\phi/u$ constant (either by increasing $J/U$ or decreasing $N$).
+  \c_{p\mu}^{(\phi)\dag} \c_{p\nu}^{(\phi)}
+  \c_{q\nu}^{(\phi)\dag} \c_{q\mu}^{(\phi)}.
+\end{align}
+However, $\H_{\t{int}}^{(\phi)}$ cannot be reduced to a spin Hamiltonian that addresses the same spin degrees of freedom as $\H_{\t{kin}}^{(\phi)}$.
+Nonetheless, for sufficiently weak SOC (small $s\phi$) the effect of $\H_{\t{int}}^{(\phi)}$ may be approximated by $\H_{\t{int}}^{(0)}$, and in turn $\H_{\t{int}}$ in Eq.~\eqref{eq:H_int_start}.
+The validity of this approximation has been previously benchmarked for SU(2)-symmetric interactions \cite{he2019engineering, smale2019observation}, and we provide additional benchmarking for SU(4) and SU(6) in Appendix \ref{sec:benchmarking}.
+To ensure that $\H_{\t{kin}}^{(\phi)}$ does not become trivial as $\phi\to0$, we can keep $J\phi/u$ constant, either by increasing $J/U$ or decreasing $N/L$.
 Altogether, the interacting spin Hamiltonian in the gauge frame becomes
@@
 \begin{align}
-  \bbk{\ss}_\MF = \f1{N^2} \times \bbk{\v\S\cdot\v\S}_\MF.
+  \bbk{\sds}_\MF = \f1{N^2} \times \bbk{\v\S\cdot\v\S}_\MF.
 \end{align}
@@
 \caption{
-Time-averaged magnetization $\sigma_\MF$ and interaction energy $\bbk{\ss}_\MF$ for different spin dimensions $n$ (indicated in the legend) as determined by mean-field simulations of $N=100$ spins initially in the x-polarized state $\ket\X$ for a time $T=10^5/u$.
+Time-averaged magnetization $\sigma_\MF$ and interaction energy $\bbk{\sds}_\MF$ for different spin dimensions $n$ (indicated in the legend) as determined by mean-field simulations of $N=100$ spins initially in the x-polarized state $\ket\X$ for a time $T=10^5/u$.
 Insets show same data after rescaling $h\to h \times (n/2)^{1/3}$, and transforming vertical axes according to Eq.~\eqref{eq:rescale}.
@@
 
-Figure \ref{fig:mean_mag_int_X} shows the time-averages of the magnetization $\sigma_\MF$ and interaction energy $\bbk{\ss}_\MF$ as computed by mean-field simulations of $N=100$ spins initially in the x-polarized state $\ket\X\equiv\ket\x^{\otimes N}$, where
+Figure \ref{fig:mean_mag_int_X} shows the time-averages of the magnetization $\sigma_\MF$ and interaction energy $\bbk{\sds}_\MF$ as computed by mean-field simulations of $N=100$ spins initially in the x-polarized state $\ket\X\equiv\ket\x^{\otimes N}$, where
 \begin{align}
@@
 As expected, the spin model exhibits a mean-field dynamical phase transition between an ordered phase at small $h$ and a disordered phase at large $h$.
-The ordered phase has a non-zero magnetization $\sigma_\MF$ and an interaction energy $\bbk{\ss}_\MF$ that asymptotically approach their maximal values as $h\to0$.
-The disordered phase has no (time-averaged) magnetization, $\sigma_\MF=0$, but the interaction energy $\bbk{\ss}_\MF$ nonetheless indicates persistent nontrivial inter-spin correlations when $n>2$.
-These nontrivial correlations vanish as $h\to\infty$, in which case $\bbk{\ss}_\MF$ approaches the minimal value allowed by conservation laws (clarified below).
+The ordered phase has a non-zero magnetization $\sigma_\MF$ and an interaction energy $\bbk{\sds}_\MF$ that asymptotically approach their maximal values as $h\to0$.
+The disordered phase has no (time-averaged) magnetization, $\sigma_\MF=0$, but the interaction energy $\bbk{\sds}_\MF$ nonetheless indicates persistent nontrivial inter-spin correlations when $n>2$.
+These nontrivial correlations vanish as $h\to\infty$, in which case $\bbk{\sds}_\MF$ approaches the minimal value allowed by conservation laws (clarified below).
 By minimizing the reduced field $h$ for which $\sigma_\MF=0$, we numerically find that the transition between ordered and disordered phases occurs at a critical field $h_\crit=\p{n/2}^{-\alpha}$ with $\alpha\approx1/3$ (see Figure \ref{fig:crit_fields_X}).
@@
   &&
-  \bbk{\ss}_\MF \to \f{\bbk{\ss}_\MF-\gamma\p{n}}{1-\gamma\p{n}},
+  \bbk{\sds}_\MF \to \f{\bbk{\sds}_\MF-\gamma\p{n}}{1-\gamma\p{n}},
   \label{eq:rescale}
@@
 The time-averaged transverse magnetization $\sigma_\MF$ then trivially vanishes as $h\to\infty$.
-The interaction energy $\bk{\ss}_\MF = \bk{\bar{\v s}}_\MF \cdot \bk{\bar{\v s}}_\MF + O(1/N)$, meanwhile, has contributions from:
+The interaction energy $\bk{\sds}_\MF = \bk{\bar{\v s}}_\MF \cdot \bk{\bar{\v s}}_\MF + O(1/N)$, meanwhile, has contributions from:
 \begin{enumerate*}
 \item the diagonal parts of the mean spin matrix $\bk{\bar{\v s}}_\MF$, which are conserved by inhomogeneous spin precession, and
-\item the off-diagonal parts of $\bk{\bar{\v s}}_\MF$, whose oscillations average to zero when evaluating the time average in $\bbk{\ss}_\MF$.
+\item the off-diagonal parts of $\bk{\bar{\v s}}_\MF$, whose oscillations average to zero when evaluating the time average in $\bbk{\sds}_\MF$.
 \end{enumerate*}
-Altogether, the interaction energy $\bbk{\ss}_\MF$ in the strong-field limit is determined by the time-independent diagonal part $\diag\bk{\bar{\v s}}_\MF = \diag\op{\x}$, namely
-\begin{align}
-  \lim_{h\to\infty} \bbk{\ss}_\MF
+Altogether, the interaction energy $\bbk{\sds}_\MF$ in the strong-field limit is determined by the time-independent diagonal part $\diag\bk{\bar{\v s}}_\MF = \diag\op{\x}$, namely
+\begin{align}
+  \lim_{h\to\infty} \bbk{\sds}_\MF
   = \Tr\sp{\p{\diag\op{\x}}^2}
@@
 where we have used the fact that the axial magnetizations $\bk{\s_{\z,q}}=\frac1N\bk{\S_\z}$ within the PS manifold, and the initial value of $\bk{\S_\z}=0$ is conserved by $\H_\spin$.
-The weak-field effective Hamiltonian preserves permutational symmetry, so $\bbk{\ss}_\MF\to1$ as $h\to0$.
+The weak-field effective Hamiltonian preserves permutational symmetry, so $\bbk{\sds}_\MF\to1$ as $h\to0$.
 Moreover, the initial y-magnetization $\bk{\S_\y}=0$ is conserved by $\H_\spin$, so long-time-averaged magnetization $\sigma_\MF$ is determined by the time-average of $\s_\x$ for a single spin:
@@
 A corollary of Figure \ref{fig:mean_mag_int_X} for the initial state $\ket\XX$.
-The inset for interaction energy $\bbk{\ss}_\MF$ in Figure \ref{fig:mean_mag_int_X} subtracts off the minimal value of $\bbk{\ss}_\MF$ and rescales to lie on the interval $[0,1]$, as prescribed in Eq.~\eqref{eq:rescale}.
+The inset for interaction energy $\bbk{\sds}_\MF$ in Figure \ref{fig:mean_mag_int_X} subtracts off the minimal value of $\bbk{\sds}_\MF$ and rescales to lie on the interval $[0,1]$, as prescribed in Eq.~\eqref{eq:rescale}.
 Here the subtracting and rescaling is identical, but with a minimal value of $2\gamma\p{n}$ rather than $\gamma\p{n}$.
@@
 \end{align}
-Similarly to Figure \ref{fig:mean_mag_int_X}, Figures \ref{fig:mean_mag_int_XX} and \ref{fig:mean_mag_int_XXI} show the time-averaged magnetization $\sigma_\MF$ and interaction energy \mbox{$\bbk{\ss}_\MF$} throughout mean-field dynamics of the initial states $\ket\XX$ and $\ket\XXI$.
+Similarly to Figure \ref{fig:mean_mag_int_X}, Figures \ref{fig:mean_mag_int_XX} and \ref{fig:mean_mag_int_XXI} show the time-averaged magnetization $\sigma_\MF$ and interaction energy \mbox{$\bbk{\sds}_\MF$} throughout mean-field dynamics of the initial states $\ket\XX$ and $\ket\XXI$.
 These figures exclude the trivial case of spin dimension $n=2$, for which $\ket\xx=\ket{-s}$ is an eigenstate of $\H_\spin$ and $\ket\xxi=e^{-\i\frac{\pi}{2}\s_\z}\ket\x\equiv\ket\y$ is spin-polarized along the y axis.
@@
 
-Unlike Figure \ref{fig:mean_mag_int_X} (for $\ket\X$), Figure \ref{fig:mean_mag_int_XX} (for $\ket\XX$) exhibits no sharp transition between distinct dynamical phases: the time-averaged magnetization $\sigma_\MF=0$ for all values of the field $h$, and the interaction energy $\bbk{\ss}_\MF$ smoothly crosses over from a maximal value of 1 to a minimal value of $2\gamma\p{n}$.
-The minimal value of $\bbk{\ss}_\MF$ approached as $h\to\infty$ can be explained with arguments identical to those in the paragraph containing Eq.~\eqref{eq:lim_ss}, which now imply that
-\begin{align}
-  \lim_{h\to\infty} \bbk{\ss}_\MF
+Unlike Figure \ref{fig:mean_mag_int_X} (for $\ket\X$), Figure \ref{fig:mean_mag_int_XX} (for $\ket\XX$) exhibits no sharp transition between distinct dynamical phases: the time-averaged magnetization $\sigma_\MF=0$ for all values of the field $h$, and the interaction energy $\bbk{\sds}_\MF$ smoothly crosses over from a maximal value of 1 to a minimal value of $2\gamma\p{n}$.
+The minimal value of $\bbk{\sds}_\MF$ approached as $h\to\infty$ can be explained with arguments identical to those in the paragraph containing Eq.~\eqref{eq:lim_ss}, which now imply that
+\begin{align}
+  \lim_{h\to\infty} \bbk{\sds}_\MF
   = \Tr\sp{\p{\diag\op{\xx}}^2}
@@
 
-Turning now to mean-field results for the initial kitten state $\ket\XXI$ in Figure \ref{fig:mean_mag_int_XXI}, we remark that the magnetization $\sigma_\MF$ and interaction energy $\bbk{\ss}_\MF$ behave identically to those for the initial spin-polarized state $\ket\X$ in Figure \ref{fig:mean_mag_int_X}.
+Turning now to mean-field results for the initial kitten state $\ket\XXI$ in Figure \ref{fig:mean_mag_int_XXI}, we remark that the magnetization $\sigma_\MF$ and interaction energy $\bbk{\sds}_\MF$ behave identically to those for the initial spin-polarized state $\ket\X$ in Figure \ref{fig:mean_mag_int_X}.
 This finding can be understood through the fact that
@@
 \end{align}
-In turn, expanding $\ss$ according to Eq.~\eqref{eq:dot_product} shows that
-\begin{align}
-  \T_\z^\theta{}^\dag \R_\z^\theta{}^\dag \, \ss \, \R_\z^\theta \T_\z^\theta
-  = \ss,
-\end{align}
-which implies that the interaction energy $\bk{\ss}$ throughout dynamics of the initial kitten state $\ket\XXI$ is the same as that of the spin-polarized state $\ket\X$.
+In turn, expanding $\sds$ according to Eq.~\eqref{eq:dot_product} shows that
+\begin{align}
+  \T_\z^\theta{}^\dag \R_\z^\theta{}^\dag \, \sds \, \R_\z^\theta \T_\z^\theta
+  = \sds,
+\end{align}
+which implies that the interaction energy $\bk{\sds}$ throughout dynamics of the initial kitten state $\ket\XXI$ is the same as that of the spin-polarized state $\ket\X$.
 
@@
 
+\section*{Acknowledgements}
+
+We thank V.~Gurarie and E.~Yuzbashyan for useful discussions on the topic.
+This work was supported by the AFOSR grants FA9550-18-1-0319, FA9550-19-1-027, by the DARPA and ARO grant W911NF-16-1-0576, ARO W911NF-19-1-0210, DOE (QSA), NSF PHY1820885, NSF JILA-PFC PHY-1734006 and NSF QLCI-2016244 grants, and by NIST.
+
 \bibliography{main.bib}
@@
 \appendix
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Numerical benchmarking of the spin model}
+\label{sec:benchmarking}
+
+In this appendix we present numerical evidence to support the validity of the spin models derived in Sections \ref{sec:spin_model} and \ref{sec:SOC}.
+Figures \ref{fig:benchmarking_SU4} and \ref{fig:benchmarking_SU6} show a set of time-averaged observables computed with numerically exact simulations of a Fermi-Hubbard model and an effective spin model, respectively, with $n=4$ (Figure \ref{fig:benchmarking_SU4}) and $n=6$ (Figure \ref{fig:benchmarking_SU6}) internal levels per spin.
+Details for these simulations are provided in the caption of Figure \ref{fig:benchmarking_SU4}.
+Our main conclusion from these figures is that the two models show remarkable agreement for the observables considered in our work.
+Note that these results are only intended to benchmark the approximation of a Fermi-Hubbard model by a spin model; these results are not expected to agree with the mean-field theory in Section \ref{sec:mean_field} due to strong finite-size effects.
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=\textwidth]{benchmarking_SU4.pdf}
+\caption{
+Numerical results (analogous to Figures \ref{fig:mean_mag_int_X}, \ref{fig:mean_mag_int_XX}, and \ref{fig:mean_mag_int_XXI} of the main text) for the time-averaged interaction energy and magnetization (both normalized to a maximal value of 1) in a system of $L=5$ lattice sites, for both a Fermi-Hubbard model (dots) and spin model (lines) with $n=4$ internal states per spin.
+The corresponding initial state (defined in Section \ref{sec:mean_field} of the main text) is indicated in each panel, and observables are averaged over a time $tJ=200$.
+Color indicates the value of $U/J$, and the field $h$ corresponds to $2J\phi/u$ in the case of the Fermi-Hubbard model.
+Simulations are performed in real-space, with spin-orbit coupling (SOC) implemented through a homogeneous drive (with no site or $\phi$ dependence) and nearest-neighbor tunneling terms that contain factors of $e^{\pm\i\mu\phi}$.
+Results for the initial kitten state $\ket\XXI$ are excluded because they are identical to those of $\ket\X$, and magnetization for the initial state $\ket\XX$ is always $0$.
+Note that while panels (a) and (b) are representative of infinite-time behavior, the inset in panel (c) shows that the Fermi-Hubbard and spin models exhibit different behaviors on very long time scales, although good agreement is restored by rescaling time in the spin model, indicating the likelihood of a need to renormalize spin model parameters.
+In any case, such time scales are inaccessible in current experiments and diverge as $N\to\infty$, so these corrections do not affect the main results of our work.
+}
+\label{fig:benchmarking_SU4}
+\end{figure}
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=\textwidth]{benchmarking_SU6.pdf}
+\caption{
+Numerical results identical to Figure \ref{fig:benchmarking_SU4}, but with $L=4$ lattice sites and $n=6$ internal states per spin.
+}
+\label{fig:benchmarking_SU6}
+\end{figure}
 
@@
 \item that the excited-state hyperfine manifold had the same total spin $s$ as the ground-state manifold, and
-\item that all drive amplitudes are real, which enforces a phase-locking condition between the driving lasers.
+\item that all drive amplitudes are real (which enforces a phase-locking condition between the driving lasers).
 \end{enumerate*}
-If the excited-state hyperfine manifold has total spin $s+r$ (with $r\in\set{+1,0,-1}$), then the spin-raising/lowering operators $\s_\pm$ that appear in Eq.~\eqref{eq:H_drive_full} (where $\s_{\v v,\pm,j}$ raises or lowers the spin of atom $j$ along axis $\v v$) become
+To derive an effective drive Hamiltonian for the general case in which the excited-state hyperfine manifold has total spin $s+r$ with $r\in\set{+1,0,-1}$, we decompose all lasers into their right- and left-circular polarization components and write the full drive Hamiltonian in the form
+\begin{align}
+  \H_{\t{drive}}^{\t{full}}
+  = \sum_{j,\v v,\sigma} \Omega_{\v v\sigma}
+  \p{e^{-\i\kappa\v v\cdot\v\ell j} \s_{\v v\sigma j}^{(r)}
+    \otimes\op{\e}{\g}_j + \t{h.c.}}
+  + \Delta \hat N_\e,
+\end{align}
+where $\Omega_{\v v\sigma}$ is the amplitude of $\sigma$-polarized light propagating along axis $\v v$, with $\sigma=+1$ and $-1$ respectively for right and left circular polarizations; and $\s_{\v v\sigma j}$ is a spin-raising/lowering operator for atom $j$ along axis $\v v$, defined by appropriately rotating the single-atom spin operators
 \begin{align}
@@
   \hat T_\pm^{(r)} \equiv \mp\sqrt{\f{2(s+r)+1}{2\ell+1}}
-  \sum_\mu \bk{s\mu;1,\pm1|s+r,\mu\pm1} \op{\mu\pm1}{\mu},
-\end{align}
-where $\bk{j_1m_1;j_2m_2|j_3m_3}$ is a Clebsch-Gordan coefficient, and we have normalized $\hat T_\pm^{(r)}$ such that $\tr\sp{\hat T_\pm^{(r)}{}^\dag \hat T_\pm^{(r)}}=1$.
-Still assuming real drive amplitudes, the corresponding effective drive Hamiltonian that replaces Eq.~\eqref{eq:H_3LD_single} is then
+  \sum_\mu \bk{s\mu;1,\pm1|s+r,\mu\pm1} \op{\mu\pm1}{\mu}.
+\end{align}
+Here $\bk{j_1m_1;j_2m_2|j_3m_3}$ is a Clebsch-Gordan coefficient, and we have normalized $\hat T_\pm^{(r)}$ such that $\tr\sp{\hat T_\pm^{(r)}{}^\dag \hat T_\pm^{(r)}}=1$.
+Still assuming real drive amplitudes, the corresponding effective drive Hamiltonian that replaces Eq.~\eqref{eq:H_3LD_single} in the far-detuned limit $\abs{\Delta}\gg\abs{\Omega_{\v v\sigma}}$ is then
 \begin{align}
@@
 \end{align}
-where $f_r^{(k)}$ are $n$-dependent scalars:
+where $f_r^{(k)}$ are scalars that depend on the spin dimension $n$:
 \begin{align}
@@
 \end{align}
-The normalized, dimensionless, $(n^2-1)$-component Lax vector $\vec\ell\p{z}$ associated with $\H_\spin$, which is defined with an auxiliary complex parameter $z$, has components
+The intensive, dimensionless, $(n^2-1)$-component Lax vector $\vec\ell\p{z}$ associated with $\H_\spin$, which is defined with an auxiliary complex parameter $z$, has components
 \begin{align}

### 2021-09-08 sun_phases/spin_model_paper/main.tex
final changes to spin model paper before blueforming

@@
 \caption{
-Drive Hamiltonians (left column) that can be implemented with different amplitude-matching conditions (right column), some of which are specified by an arbitrary sign $\sigma\in\set{+1,-1}$.
-The drives shown here are equal to that Eq.~\eqref{eq:H_3LD_single} up to a possible energy shift of $\s_\x^2+\s_\y^2+\s_\z^2=s(s+1)$, and come in mutually commuting pairs: a drive with $\Omega_m=1$ and $\Omega_n=0$ for both $n\ne m$ commutes with the drive in which $\Omega_m=0$ and both $\abs*{\Omega_n}=1$.
+Drive Hamiltonians (left column) that can be implemented with different amplitude-matching conditions (right three columns), some of which are specified by an arbitrary sign $\sigma\in\set{+1,-1}$.
+The drives shown here are equal to that Eq.~\eqref{eq:H_3LD_single} up to a possible energy shift of $\s_\x^2+\s_\y^2+\s_\z^2=s(s+1)$, and come in mutually commuting pairs: a drive with $\abs*{\Omega_m}=1$ and $\Omega_n=0$ for both $n\ne m$ commutes with the drive in which $\Omega_m=0$ and both $\abs*{\Omega_n}=1$.
 }
 \vspace{.5em}
-\begin{tabular}{c|c}
+\begin{tabular}{c||c|c|c}
   $\H_{\t{drive}}^{\t{single}}$
-  & $(\tilde\Omega_0,\tilde\Omega_+,\tilde\Omega_-)$
+  & $\tilde\Omega_0$ & $\tilde\Omega_+$ & $\tilde\Omega_-$
   \\ \hline\hline
-  $-\s_\z^2$ & $\p{1,0,0}$
+  $-\s_\z^2$ & 1 & 0 & 0
   \\ \hline
-   $-\s_\x^2$ & $\p{0,1,0}$
+   $-\s_\x^2$ & 0 & 1 & 0
   \\ \hline
-  $-\s_\y^2$ & $\p{0,0,1}$
+  $-\s_\y^2$ & 0 & 0 & 1
   \\ \hline
-  $\sigma \s_\z + \s_\z^2$ & $\p{0,1,\sigma}$
+  $\sigma \s_\z + \s_\z^2$ & 0 & 1 & $\sigma$
   \\ \hline
-  $\sigma \s_\x + \s_\x^2$ & $\p{1,0,\sigma}$
+  $\sigma \s_\x + \s_\x^2$ & 1 & 0 & $\sigma$
   \\ \hline
-  $\sigma\p{\s_\z \s_\x+\s_\x \s_\z} + \s_\y^2$ & $\p{1,\sigma,0}$
+  $\sigma\p{\s_\z \s_\x+\s_\x \s_\z} + \s_\y^2$ & 1 & $\sigma$ & 0
   \\ \hline
   $\pm \s_\z \pm \sigma \s_\x + \sigma (\s_\z \s_\x + \s_\x \s_\z)$
-  & $\p{1,\sigma,\pm\sigma}$
+  & 1 & $\sigma$ & $\pm\sigma$
 \end{tabular}
@@
 
-In principle, making the gauge transformation $\c_{q\mu}^\dag\to \c_{q-\mu\phi,\mu}^\dag$ also transforms the interaction Hamiltonian as $\H_{\t{int}} \to \H_{\t{int}}^{(\phi)}$, where
-\begin{align}
-  \H_{\t{int}}^{(\phi)}
-  \equiv \f{u}{2N} \sum_{k,\ell,p,q,\mu,\nu}
-  \c_{k\mu}^{(\phi)\dag} \c_{\ell\mu}^{(\phi)}
-  \c_{p\nu}^{(\phi)\dag} \c_{q\nu}^{(\phi)}
-  \times \delta_{k+p,\ell+q},
-\end{align}
-and $\c_{k\mu}^{(\phi)}\equiv\c_{k-\mu\phi,\mu}$.
-However, $\H_{\t{int}}^{(\phi)}$ cannot be reduced to a spin Hamiltonian that addresses the same spin degrees of freedom as $\H_{\t{kin}}^{(\phi)}$.
-Again making a frozen-mode approximation, or equivalently keeping only terms that preserve the coherences of an initially spin-polarized state, the gauge-transformed interaction Hamiltonian becomes
-\begin{align}
-  \H_{\t{int}}^{(\phi)}
-  \equiv -\f{u}{2N} \sum_{p,q,\mu,\nu}
-  \c_{p\mu}^{(\phi)\dag} \c_{p\nu}^{(\phi)}
-  \c_{q\nu}^{(\phi)\dag} \c_{q\mu}^{(\phi)}.
-\end{align}
-However, $\H_{\t{int}}^{(\phi)}$ cannot be reduced to a spin Hamiltonian that addresses the same spin degrees of freedom as $\H_{\t{kin}}^{(\phi)}$.
-Nonetheless, for sufficiently weak SOC (small $s\phi$) the effect of $\H_{\t{int}}^{(\phi)}$ may be approximated by $\H_{\t{int}}^{(0)}$, and in turn $\H_{\t{int}}$ in Eq.~\eqref{eq:H_int_start}.
-The validity of this approximation has been previously benchmarked for SU(2)-symmetric interactions \cite{he2019engineering, smale2019observation}, and we provide additional benchmarking for SU(4) and SU(6) in Appendix \ref{sec:benchmarking}.
+Finally, the gauge transformation $\c_{q\mu}^\dag\to \c_{q-\mu\phi,\mu}^\dag$ also transforms the interaction Hamiltonian.
+Applying this transformation to Eq.~\eqref{eq:H_int_momenta} and keeping only terms that respect the coherences imposed on initial states by the laser drive, we can again derive an effective spin model.
+For sufficiently weak SOC ($s\phi\to0$) this spin model is still well-approximated by $\H_{\t{int}}$ in Eqs.~\eqref{eq:H_int_start} and \eqref{eq:H_int}.
+The validity of this approximation has been previously benchmarked for SU(2)-symmetric interactions \cite{he2019engineering, smale2019observation}, and we provide additional benchmarking for SU(4) and SU(6) in Appendix \ref{sec:benchmarking} (which finds that the spin model works well even for {\it large} $\phi$).
 To ensure that $\H_{\t{kin}}^{(\phi)}$ does not become trivial as $\phi\to0$, we can keep $J\phi/u$ constant, either by increasing $J/U$ or decreasing $N/L$.
@@
 \end{align}
-and $\ket{-\x}$ is a state polarized along $-$x, defined similarly to $\ket\x$ in Eq.~\eqref{eq:state_x}:
+and $\ket{-\x}$ is a state polarized along $-x$, defined similarly to $\ket\x$ in Eq.~\eqref{eq:state_x}:
 \begin{align}
@@
 Similarly to Figure \ref{fig:mean_mag_int_X}, Figures \ref{fig:mean_mag_int_XX} and \ref{fig:mean_mag_int_XXI} show the time-averaged magnetization $\sigma_\MF$ and interaction energy \mbox{$\bbk{\sds}_\MF$} throughout mean-field dynamics of the initial states $\ket\XX$ and $\ket\XXI$.
-These figures exclude the trivial case of spin dimension $n=2$, for which $\ket\xx=\ket{-s}$ is an eigenstate of $\H_\spin$ and $\ket\xxi=e^{-\i\frac{\pi}{2}\s_\z}\ket\x\equiv\ket\y$ is spin-polarized along the y axis.
+These figures exclude the trivial case of spin dimension $n=2$, for which $\ket\xx=\ket{-s}$ is an eigenstate of $\H_\spin$ and $\ket\xxi=e^{-\i\frac{\pi}{2}\s_\z}\ket\x\equiv\ket\y$ is spin-polarized along the $y$ axis.
 The first and perhaps most interesting observation to make about Figures \ref{fig:mean_mag_int_XX} and \ref{fig:mean_mag_int_XXI} is that they are different, signifying the importance of intra-spin coherences for the dynamical behavior of multilevel spin models.

### 2021-09-08 sun_phases/qudit_tomo.tex
rename tomography paper

--- 2021-08-30 sun_phases/spin_qudit_tomo.tex
@@
 In Section \ref{sec:polarization_ops}, we introduce a set of qudit operators that are closely related to the spherical harmonics, and which play a central role in our work.
-We then map the quantum problem of spin qudit tomography onto the classical problem of signal recovery on the sphere in Section \ref{sec:signal_recovery}, thereby providing an interesting new perspective on spin qudit tomography.
+We then map the quantum problem of spin qudit tomography onto the classical problem of signal recovery on the sphere in Section \ref{sec:signal_recovery}, thereby providing an intuitive perspective on spin qudit tomography.
 In Section \ref{sec:error} we provide a priori upper bounds and a posteriori estimates of the statistical error in a qudit state reconstructed from measurements of spin projection along a given set of $r$ measurement axes.
@@
 This representation is faithful in the sense that $X$ is uniquely determined by the phase-space values $X^{\t{PS}}\p{\bm v}$ at all points $\bm v$ on the sphere.
-The transformation rules for polarization operators in Eq.~\eqref{eq:trans_rot}, together with the fact that $\bk{s|T_{\ell m}|s}=0$ unless $m=0$, suffice to show that (see Appendix \ref{sec:rotations})
+The transformation rules for polarization operators in Eq.~\eqref{eq:trans_rot}, together with the fact that $\bk{s|T_{\ell m}|s}=0$ unless $m=0$, suffice to show that
 \begin{align}
@@
 \end{align}
-where the scalar $c_\ell$ simply enforces $\obk{T_{\ell m}|T_{\ell m}}=1$.
+where the scalar $c_\ell$ simply enforces $\obk{T_{\ell m}|T_{\ell m}}=1$ (see Appendix \ref{sec:rotations}).
 The polarization operators $T_{\ell m}$ are thus a quantum analogue of the spherical harmonics $Y_{\ell m}$, and play an important role in phase-space formalisms for spin qudits \cite{li2013weylwignermoyal}.
@@
 Here $\bm v$ and $\ell$ label a row of $M_V$, or equivalently label a standard (``one-hot'') basis vector $\ket{\bm v\ell}$ of a $\p{\abs{V}\times d}$-dimensional vector space, and $\obra{T_{\bm v\ell,0}}$ is the conjugate transpose of $\oket{T_{\bm v\ell,0}}$.
-A necessary and sufficient condition for $V$ to allow for full state tomography is that the measured polarization operators $T_{\bm v\ell,0}$, or equivalently the columns of $M_V$, span the entire ($d^2$-dimensional) space of operators on a $d$-level spin qudit.
+A necessary and sufficient condition for $V$ to allow for full state tomography is that the measured polarization operators $T_{\bm v\ell,0}$, or equivalently the rows of $M_V$, span the entire ($d^2$-dimensional) space of operators on a $d$-level spin qudit.
 In this case $M_V$ must be full rank, with $d^2$ nonzero singular values.
@@
 \end{align}
-where, using the fact that $T_j=T_j^\dag$ (because they are diagonal, degree $m=0$ polarization operators),
+where, using the fact that $T_j=T_j^\dag$ (because they are diagonal polarization operators with degree $m=0$),
 \begin{align}
@@
 The assumption that observables can be estimated up to uncorrelated noise with maximal variance $\epsilon^2$, summarized by Eq.~\eqref{eq:error_assumption}, is reasonable when measurement error is dominated by classical sources of experimental noise.
-However, this assumption breaks down when measurement error is limited by fundamental quantum (shot) noise.
+However, this assumption breaks down when measurement error is limited by fundamental quantum shot noise (i.e.~finite sampling error).
 We relax the assumption of Eq.~\eqref{eq:error_assumption} in Appendix \ref{sec:bound}, where we instead assume that $\tilde\rho_V$ is built from $n$ independent measurements of spin projection along every axis $\bm v\in V$, with shot noise the dominant source of error.
@@
 \end{align}
-where the {\it quantum error scale} $\epsilon_V$, so called because the bound in Eq.~\eqref{eq:bound} is set by quantum shot noise, is defined in terms of the spectral range of $T_{\ell,0}$:
+where the {\it quantum error scale} $\epsilon_V$ is defined in terms of the spectral range of $T_{\ell,0}$:
 \begin{align}
@@
 The individual bounds on $\bbk{\epsilon_{\bm v\ell}^2}$ for each axis $\bm v$ and degree $\ell$ are tight, but these bounds cannot all be achieved simultaneously.
-There is therefore still room for improvement on the bound in Eq.~\eqref{eq:bound} by maximizing $\E_V$ over the set of all qudit states $\rho$.
+There is therefore still room for improvement on the bound in Eq.~\eqref{eq:bound} by maximizing $\E_V$ over the set of all physical qudit states $\rho$.
 We discuss this maximization problem in Appendix \ref{sec:exact}, but leave its full solution to future work.
-We also note that the (squared) reconstruction error bound in Eq.~\eqref{eq:bound} obeys the ``standard quantum limit'' of $\sim1/n$ scaling in the number of measurements.
+We also note that the reconstruction error bound in Eq.~\eqref{eq:bound} obeys the ``standard quantum limit'' of $\sim1/n$ scaling in the number of measurements.
 In principle, this scaling can be improved to $\sim1/n^2$ by preparing and measuring entangled copies of many qudits \cite{giovannetti2006quantum}.
@@
 \item For each axis $\bm v\in V$, make $n$ measurements of spin projection, and set $\tilde\Pi_{\bm v\mu}\approx\bk{\Pi_{\bm v\mu}}_\rho$ to the fraction of times in which the measurement outcome was $\mu$.
-\item Use the the estimates $\tilde\Pi_{\bm v\mu}$ of $\bk{\Pi_{\bm v\mu}}_\rho$ to compute the estimates
+\item Use the the estimates $\tilde\Pi_{\bm v\mu}$ of $\bk{\Pi_{\bm v\mu}}_\rho$ to compute estimates of $\bk{T_{\bm v\ell,0}}_\rho$,
 \begin{align}
   \tilde T_{\bm v\ell,0}
-  \equiv \sum_\mu \bk{\mu|T_{\ell,0}|\mu} \tilde\Pi_{\bm v\mu}
-\end{align}
-of $\bk{T_{\bm v\ell,0}}_\rho$, where the matrix elements of $T_{\ell,0}$ are provided in Eq.~\eqref{eq:trans_op}.
+  \equiv \sum_\mu \bk{\mu|T_{\ell,0}|\mu} \tilde\Pi_{\bm v\mu},
+\end{align}
+where the matrix elements of $T_{\ell,0}$ are provided in Eq.~\eqref{eq:trans_op}.
 \item Denoting the nonzero singular values of $M_{V\ell}$ by $M^V_{\ell k}$ and the corresponding left singular vectors by $\bm x^V_{\ell k} = \sum_{\bm v} x^V_{\ell k \bm v} \ket{\bm v}$, compute the  operators and coefficients
@@
 For reference, Figure \ref{fig:axes} also shows the smallest measurement-adjusted error scales $\beta_\theta$ achievable with the method of Ref.~\cite{newton1968measurability}, which is comparable to those achieved with our randomized protocol at $\abs{V}\approx3d$.
-The method of Ref.~\cite{newton1968measurability} requires choosing an angle $\theta$, and provides no prescription for making this choice.
+The method of Ref.~\cite{newton1968measurability} requires choosing an angle $\theta$, namely the polar angle of all measurement axes, and provides no prescription for making this choice.
 We therefore find the optimal choice of $\theta$ by minimizing the error scale $\beta_\theta$ over all $\theta$ (see Appendix \ref{sec:angle_search}), and show $\min\beta_\theta/\tilde\beta(0)$ in Figure \ref{fig:axes}.
-Empirically, we find that the optimal angle for the method of Ref.~\cite{newton1968measurability} is $\theta_{\t{opt}}\approx\frac{\pi}{2}(1 - \frac1{1.34d})$ (see Appendix \ref{sec:angle_search}), although we note that a randomized tomography protocol is likely more robust to errors in axis orientation because measurement axes are generally more spaced out.
-We leave the question of robustness to errors in axis orientation and the effect of these errors on state reconstruction error to future work.
+Empirically, we find that the optimal angle for the method of Ref.~\cite{newton1968measurability} is $\theta_{\t{opt}}\approx\frac{\pi}{2}(1-\frac1{1.34d})$ (see Appendix \ref{sec:angle_search}), which approaches $\pi/2$ as $d\to\infty$.
+However, the error scale $\beta_{\pi/2}=\infty$, reflecting the fact that full state tomography is impossible with measurement axes lying in a single plane.
+The method of Ref.~\cite{newton1968measurability} therefore requires extremely careful fine-tuning of measurement axis orientations for large spin dimensions.
+For this reason, we expect our randomized tomography protocol be be more robust to errors in axis orientation.
+We leave a detailed analysis of robustness to errors in axis orientation and the effect of these errors on state reconstruction to future work.
 
 As a final point, we note that any information about an unknown qudit state $\rho$, obtained from prior knowledge or preliminary measurement data, can be used to construct tailored or adaptive measurement protocols \cite{huszar2012adaptive, ferrie2014selfguided, granade2016practical, pereira2018adaptive} that are more efficient in terms of the number of measurements required to estimate $\rho$ to a fixed precision.
-We leave the development of tailored and adaptive measurement protocols to future work.
+We leave the development of tailored and adaptive measurement protocols to future work as well.
 
@@
 \onecolumngrid
-\newpage
 \appendix

### 2021-09-30 sun_phases/qudit_tomo.tex
added reference and fixed bibliography in qudit tomography paper
updating submitted versions of spin model and tomography papers

@@
 Due to advancements in experimental capabilities to address nuclear spin states (i.e.~hyperfine levels) in ultracold atomic systems \cite{daley2011quantum, lu2011strongly, mischuck2012control, aikawa2012boseeinstein, smith2013quantum, cazalilla2014ultracold}, as well as developments in the control of ultracold molecular systems \cite{bohn2017cold, takekoshi2014ultracold, zeppenfeld2012sisyphus, kozyryev2017sisyphus, puri2017synthesis, wu2017cryofuge, marco2019degenerate, liu2019molecular, anderegg2019optical, chou2020frequencycomb, lin2020quantum}, a particular setting of growing interest is the spin qudit, or a multilevel quantum angular momentum degree of freedom.
-Spin qudits can provide advantages over their qubit counterparts for quantum sensing \cite{hemmer2018squeezing, evrard2019enhanced}, enable quantum simulations of SU($d$) magnetism \cite{cazalilla2014ultracold, banerjee2013atomic, zhang2014spectroscopic, scazza2014observation, goban2018emergence}, % todo: cite SU(n) spin model paper!
-and offer unique capabilities for quantum error correction \cite{albert2020robust, gross2021designing}.
+Spin qudits can provide advantages over their qubit counterparts for quantum sensing \cite{hemmer2018squeezing, evrard2019enhanced}, enable quantum simulations of SU($d$) magnetism \cite{cazalilla2014ultracold, banerjee2013atomic, zhang2014spectroscopic, scazza2014observation, goban2018emergence, perlin2021engineering}, and offer unique capabilities for quantum computation and error correction \cite{albert2020robust, gross2021designing, barnes2021assembly}.
 In all cases, quantum state tomography is necessary to take full advantage of a spin qudit\footnote{Note that the measurement of collective observables for quantum sensing or simulation can be recast as a single-spin tomography task.}.
@@
   \caption{Signal recovery on the sphere is the problem of reconstructing an unknown function $f$ (red distribution) from its values $f\p{\bm v}$ at specific points $\bm v\in V$ (blue dots) on the sphere.
-    For most choices of $V$, reconstruction of $f$ is possible if there are at least as many points in $V$ as there are degrees of freedom in $f$.}
+    For almost all choices of $V$, reconstruction of $f$ is possible if there are at least as many points in $V$ as there are degrees of freedom in $f$.}
   \label{fig:sphere_points}
@@
 
-\bibliography{spin_qudit_tomo.bib}
+\bibliography{qudit_tomo.bib}
 \onecolumngrid

### 2021-09-30 sun_phases/spin_model_paper/main.tex
updating submitted versions of spin model and tomography papers

@@
 
-\date{\today}
+\date{22 September 2021}
 
@@
 Moreover, for $n>2$ we find that dynamical behavior can be highly sensitive to initial intra-spin coherences.
-Our predictions are readily testable in current experiments with ultracold alkalike-earth(-like) atoms.
+Our predictions are readily testable in current experiments with ultracold alkaline-earth(-like) atoms.
 \end{abstract}
@@
 Extensions of Yang-Mills and SU($n$) symmetry feature in the most well-studied examples of holographic duality \cite{maldacena1999largen} and the connection between entanglement and gravity \cite{ryu2006holographic} through the anti-de Sitter/conformal field theory (AdS/CFT) correspondence.
-In a condensed matter setting, SU($2$) appears ubiquitously as a symmetry of the Hubbard model, with important consequences for the study of quantum magnetism and high temperature superconductivity \cite{lee2006doping}.
-The extension of SU($2$) Hubbard and spin models to SU($n$) has led to predictions of exotic phases of matter such as valence bond solids \cite{read1989valencebond, rokhsar1990quadratic, kaul2012lattice, hermele2011topological} and chiral spin liquids \cite{hermele2009mott, hermele2011topological, chen2016syntheticgaugefield, nataf2016chiral}, as well as the potential to perform universal topological quantum computation \cite{freedman2004class, nayak2008nonabelian} and other phenomena \cite{nataf2014exact, nataf2016exact}.
+In a condensed matter setting, SU(2) appears ubiquitously as a symmetry of the Hubbard model, with important consequences for the study of quantum magnetism and high temperature superconductivity \cite{lee2006doping}.
+The extension of SU(2) Hubbard and spin models to SU($n$) has led to predictions of exotic phases of matter such as valence bond solids \cite{read1989valencebond, rokhsar1990quadratic, kaul2012lattice, hermele2011topological} and chiral spin liquids \cite{hermele2009mott, hermele2011topological, chen2016syntheticgaugefield, nataf2016chiral}, as well as the potential to perform universal topological quantum computation \cite{freedman2004class, nayak2008nonabelian} and other phenomena \cite{nataf2014exact, nataf2016exact}.
 Furthermore, disordered SU($n$) spin models have opened analytically tractable avenues for studying quantum chaos and information scrambling \cite{sachdev1993gapless}.
@@
 In the spirit of quantum simulation, further investigations in controlled settings will play an important role in understanding the consequence of SU($n$) symmetries for fundamental questions in physics, as well as their practical use in technological applications.
-For example, SU($2$)-symmetric spin interactions can be harnessed to develop quantum sensors that surpass classical limits on measurement precision \cite{he2019engineering, perlin2020spin}.
+For example, SU(2)-symmetric spin interactions can be harnessed to develop quantum sensors that surpass classical limits on measurement precision \cite{he2019engineering, perlin2020spin}.
 The prospect of similarly exploiting more general SU($n$) symmetries to achieve a technological advantage is still an unexplored avenue of research with untapped potential.
 
-In this work, we consider an experimentally relevant and theoretically tractable regime of the SU($n$) Hubbard model, highlighting differences and similarities with the more familiar case of SU($2$).
+In this work, we consider an experimentally relevant and theoretically tractable regime of the SU($n$) Hubbard model, highlighting differences and similarities with the more familiar case of SU(2).
 Working at ultracold temperatures and unit spatial filling (one atom per lattice site), we begin by mapping the SU($n$) Hubbard model onto a multilevel spin model with all-to-all SU($n$)-symmetric interactions in Section \ref{sec:spin_model}.
@@
 \begin{enumerate*}
-\item long-time-averaged observables obey simple scaling relations with the spin dimension $n$, exhibiting (for spin-polarized initial states) dynamical ferromagnetic and dynamical paramegnetic phases, as previously seen for the case of $n=2$ \cite{smale2019observation, lewis-swan2021cavityqed}, and
+\item long-time-averaged observables obey simple scaling relations with the spin dimension $n$, exhibiting (for spin-polarized initial states) dynamical ferromagnetic and dynamical paramagnetic phases, as previously seen for the case of $n=2$ \cite{smale2019observation, lewis-swan2021cavityqed}, and
 \item for $n>2$ the long-time dynamics can be highly sensitive to the intra-spin coherences of the initial state.
@@
 In the present work, we neglect inter-site interactions and interaction-assisted hopping, which is a good approximation for a sufficiently deep lattice, namely when $J\lesssim E_{\t{R}}$, where $E_{\t{R}}$ is the atom recoil energy.
-For simplicity, we now assume a one-dimensional periodic lattice of $L$ sites, and expand the on-site fermionic operators in terms of operators addressing (quasi-)momentum modes $q$ (in units with lattice spacing $a=1$), $\c_{j\mu} = \frac1{\sqrt{L}} \sum_k e^{-\i q\cdot j} \c_{q\mu}$, finding that
+For simplicity, we now assume a one-dimensional periodic lattice of $L$ sites, and expand the on-site fermionic operators in terms of operators addressing (quasi-)momentum modes $q$ (in units with lattice spacing $a=1$), $\c_{j\mu} = \frac1{\sqrt{L}} \sum_q e^{-\i q\cdot j} \c_{q\mu}$, finding that
 \begin{align}
@@
 
-If the interaction energy $U$ is small compared to the single-particle bandwidth $4J$, then the mode-changing collisions in $\H_{\t{int}}$ become off-resonant, motivating the frozen-mode approximation $\set{k,p}=\set{\ell,q}$ (which we assume for the remainder of this work)\footnote{Note that the frozen-mode approximation neglects correlated momentum-hopping terms of the form $\c_{\pi-p,\mu}^\dag \c_{\pi-q,\mu} \c_{p\nu}^\dag \c_{q\nu}$, which conserve both momentum and energy.
+If the interaction energy $U$ is small compared to the single-particle bandwidth $4J$, then the mode-changing collisions in $\H_{\t{int}}$ become off-resonant, motivating the frozen-mode approximation $\set{k,p}=\set{\ell,q}$ (i.e.~either $k=\ell$ and $p=q$, or $k=q$ and $p=\ell$)\footnote{Note that the frozen-mode approximation neglects correlated momentum-hopping terms of the form $\c_{\pi-p,\mu}^\dag \c_{\pi-q,\mu} \c_{p\nu}^\dag \c_{q\nu}$, which conserve both momentum and energy.
 We defer a careful treatment of these terms to future work, noting only that they vanish on the manifold of permutationally symmetric spin states with one atom per lattice site, and that the frozen-mode approximation is benchmarked in Refs.~\cite{he2019engineering, smale2019observation} and Appendix \ref{sec:benchmarking}.}.
-The terms with $k=\ell$ and $p=q$ are $\frac{u}{2N} \sum \c_{\ell\mu}^\dag \c_{\ell\mu} \c_{q\nu}^\dag \c_{q\nu} = \frac12 N u$, which is a constant  energy shift that we can freely neglect.
+The terms with $k=\ell$ and $p=q$ are $\frac{u}{2N} \sum \c_{k\mu}^\dag \c_{k\mu} \c_{p\nu}^\dag \c_{p\nu} = \frac12 N u$, which is a constant  energy shift that we can freely neglect.
 Defining the spin operators $\s_{\mu\nu q}\equiv \c_{q\mu}^\dag \c_{q\nu}$, the remaining terms of the kinetic and interaction Hamiltonians are
@@
 
-To further simplify the interaction Hamiltonian $\H_{\t{int}}$ and write it in a form reminiscent of more familiar SU($2$) spin models, we now construct the operator-valued spin matrix
+To further simplify the interaction Hamiltonian $\H_{\t{int}}$ and write it in a form reminiscent of more familiar SU(2) spin models, we now construct the operator-valued spin matrix
 \begin{align}
@@
 \end{align}
-where $\v\S\equiv\sum_q\v\s_q$ is a collective spin matrix, analogous to the collective spin vector $\vec S=(\S_\x,\S_\y,\S_\z)$ in the case of SU($2$) \cite{he2019engineering}, with $\frac12\v\S\cdot\v\S \simeq \vec S\cdot\vec S = \S_\x^2 + \S_\y^2 + \S_\z^2$ when $n=2$ (here $\simeq$ denotes equality up to identity terms).
+where $\v\S\equiv\sum_q\v\s_q$ is a collective spin matrix, analogous to the collective spin vector $\vec S=(\S_\x,\S_\y,\S_\z)$ in the case of SU(2) \cite{he2019engineering}, with $\frac12\v\S\cdot\v\S \simeq \vec S\cdot\vec S = \S_\x^2 + \S_\y^2 + \S_\z^2$ when $n=2$ (here $\simeq$ denotes equality up to identity terms).
 
@@
 (a) Ultracold atoms on a lattice of $L$ sites tunnel between neighboring lattice sites at a rate $J$, and locally repel each other with interaction energy $U$.
-(b,c) When the interaction energy $U$ is small compared to the single-particle bandwidth $4J$, the frozen-mode approximation enables the interaction Hamiltonian to be written as a spin model consisting of exchange terms $\v\s_p\cdot\v\s_q$, which swap the states two spins pinned to momentum modes $p,q$.
+(b,c) When the interaction energy $U$ is small compared to the single-particle bandwidth $4J$, the frozen-mode approximation enables the interaction Hamiltonian to be written as a spin model consisting of exchange terms $\v\s_p\cdot\v\s_q$, which swap the states of two spins pinned to modes $p,q$.
 (d) Interactions open an energy gap $u=U\times N/L$ between the manifold of permutationally symmetric states of $N$ spins, and the orthogonal complement of states that break spin-permutation symmetry.
@@
 The term $-\v\s_p\cdot\v\s_q$ thereby assigns a definite energy of $-1$ ($+1$) to a pair of spins that are symmetric (anti-symmetric) under exchange.
-In this sense, $\v\s_p\cdot\v\s_q$ is analogous to the enforcement of SU($2$) spin alignment by ferromagnetic interactions, which similarly assigns different energies to the anti-symmetric spin-0 singlet $\ket{\up\dn}-\ket{\dn\up}$ and the symmetric spin-1 triplets $\set{\ket{\up\up},\ket{\dn\dn},\ket{\up\dn}+\ket{\dn\up}}$.
-By summing over all pair-wise exchange terms $-\v\s_p\cdot\v\s_q$, the interaction Hamiltonian $\H_{\t{int}}$ energetically enforces a permutational symmetry among all spins, opening an energy gap $u$ between the manifold of all permutationally symmetric (PS) states and the orthogonal complement of excited (e.g.~spin-wave) states that break permutational symmetry.
+In this sense, $\v\s_p\cdot\v\s_q$ is analogous to the enforcement of SU(2) spin alignment by ferromagnetic interactions, which similarly assigns different energies to the anti-symmetric spin-0 singlet $\ket{\up\dn}-\ket{\dn\up}$ and the symmetric spin-1 triplets $\set{\ket{\up\up},\ket{\dn\dn},\ket{\up\dn}+\ket{\dn\up}}$.
+By summing over all pair-wise exchange terms $\v\s_p\cdot\v\s_q$, the interaction Hamiltonian $\H_{\t{int}}$ energetically enforces a permutational symmetry among all spins, opening an energy gap $u$ between the manifold of all permutationally symmetric (PS) states and the orthogonal complement of excited (e.g.~spin-wave) states that break permutational symmetry.
 See Figure \ref{fig:spin_model} for a summary of this section thus far.
 
-In the case of SU($2$), the PS manifold is precisely the Dicke manifold of collective states $\ket{m_\z}$ with total spin $S=\frac{N}{2}$ and definite spin projection $m_\z\in\set{S,S-1,\cdots,-S}$ onto a fixed quantization axis.
+In the case of SU(2), the PS manifold is precisely the Dicke manifold of collective states $\ket{m_\z}$ with total spin $S=\frac{N}{2}$ and definite spin projection $m_\z\in\set{S,S-1,\cdots,-S}$ onto a fixed quantization axis.
 Equivalently, Dicke states $\ket{m_\z}=\ket{m_\up,m_\dn}$ can be labeled by a definite number of spins $m_\up=S+m_\z$ ($m_\dn=S-m_\z$) pointing up (down) along the spin quantization axis, with $m_\up+m_\dn=N$.
@@
 \caption{
-Whereas the state of a two-level spin (qubit) can be represented by a point on (or inside) the Bloch sphere, the state of a $n$-level spin is more generally represented by a probability distribution on the sphere.
+Whereas the state of a two-level spin (qubit) can be represented by a point on (or inside) the Bloch sphere, the state of an $n$-level spin is more generally represented by a probability distribution on the Bloch sphere.
 The distribution shown for $n=10$ corresponds to a Haar-random pure state.
@@
 The state of a two-level spin, or a qubit, is commonly represented by a point on (or within) the Bloch sphere.
-More generally, the state $\ket\psi$ of a $n$-level spin can be represented by a quasi-probability distribution $Q_\psi$ on the sphere (commonly known as the Husimi-$Q$ function, e.g.~in the spin-squeezing community \cite{ma2011quantum}).
-The value $Q_\psi\p{\v v}$ at a point $\v v$ on the sphere is equal to the overlap of $\ket\psi$ with a pure state $\ket{\v v}$ maximally polarized in the direction of $\v v$: $Q_\psi\equiv\abs{\bk{\v v|\psi}}^2$ (see Figure \ref{fig:spin_dist}).
-In the case of a mixed state $\hat\rho$, this distribution is defined by $Q_{\hat\rho}\equiv\bk{\v v|\hat\rho|\v v}$.
+More generally, the state $\ket\psi$ of an $n$-level spin can be represented by a quasi-probability distribution $Q_\psi$ on the Bloch sphere (commonly known as the Husimi-$Q$ function, e.g.~in the spin-squeezing community \cite{ma2011quantum}).
+The value $Q_\psi\p{\v v}$ at a point $\v v$ on the sphere is equal to the overlap of $\ket\psi$ with a pure state $\ket{\v v}$ that is maximally polarized in the direction of $\v v$: $Q_\psi\p{\v v}\equiv\abs{\bk{\v v|\psi}}^2$ (see Figure \ref{fig:spin_dist}).
+In the case of a mixed state $\hat\rho$, this distribution is defined by $Q_{\hat\rho}\p{\v v}\equiv\bk{\v v|\hat\rho|\v v}$.
 Closely related spherical representations of multilevel spin states and operators are discussed in Refs.~\cite{dowling1994wigner, li2013weylwignermoyal}.
@@
 \begin{enumerate*}
-\item two counter-propagating right-circularly polarized lasers with with drive amplitudes $\Omega_\pm$ and wavevectors $\kappa\v v_\pm$, propagating in opposite directions along the $z$ axis, $\v v_\pm=(0,0,\pm1)$, and
+\item two counter-propagating right-circularly polarized lasers with drive amplitudes $\Omega_\pm$ and wavevectors $\kappa\v v_\pm$, propagating in opposite directions along the $z$ axis, $\v v_\pm=(0,0,\pm1)$, and
 \item a third laser linearly polarized along $z$, with drive amplitude $\Omega_0$ and wavevector $\kappa \v v_0$, propagating along the $x$ axis $\v v_0=(1,0,0)$.
@@
 \end{align}
-where $m\in\set{+1,0,-1}$ indexes the laser pointing along $\v v_m$; the SOC angle $\phi\equiv\kappa\v v_+\cdot\v\ell=\kappa a \cos\theta$ (with $a$ the lattice spacing); $\s_{\z,j},\s_{+,j}\s_{-,j}$ are standard axial, spin-raising, and spin-lowering operators for spin $j$; $\s_{0,j}\equiv\s_{\z,j}$ for shorthand; $\ket\g_j$ and $\ket\e_j$ respectively denote the ground and excited electronic states of atom $j$; and $\hat N_\e=\1\otimes\sum_j\op\e_j$ counts the number of excited atoms (with $\1$ the identity operator on all spin degrees of freedom).
+where $m\in\set{+1,0,-1}$ indexes the laser pointing along $\v v_m$; the SOC angle $\phi\equiv\kappa\v v_+\cdot\v\ell=\kappa \cos\theta$ (in units with lattice spacing $a=1$); $\s_{\z,j},\s_{+,j}\s_{-,j}$ are standard axial, spin-raising, and spin-lowering operators for the spin at lattice site $j$; $\s_{0,j}\equiv\s_{\z,j}$ for shorthand; $\ket\g_j$ and $\ket\e_j$ respectively denote the ground and excited electronic states of atom $j$; and $\hat N_\e=\1\otimes\sum_j\op\e_j$ counts the number of excited atoms (with $\1$ the identity operator on all spin degrees of freedom).
 
@@
 Drive Hamiltonians (left column) that can be implemented with different amplitude-matching conditions (right three columns), some of which are specified by an arbitrary sign $\sigma\in\set{+1,-1}$.
-The drives shown here are equal to that Eq.~\eqref{eq:H_3LD_single} up to a possible energy shift of $\s_\x^2+\s_\y^2+\s_\z^2=s(s+1)$, and come in mutually commuting pairs: a drive with $\abs*{\Omega_m}=1$ and $\Omega_n=0$ for both $n\ne m$ commutes with the drive in which $\Omega_m=0$ and both $\abs*{\Omega_n}=1$.
+The drives shown here are equal to that of Eq.~\eqref{eq:H_3LD_single} up to a possible energy shift of $\s_\x^2+\s_\y^2+\s_\z^2=s(s+1)$, and come in mutually commuting pairs: a drive with $\abs*{\Omega_m}=1$ and $\Omega_n=0$ for both $n\ne m$ commutes with the drive in which $\Omega_m=0$ and both $\abs*{\Omega_n}=1$.
 }
@@
 Second, each of $\tilde\Omega_0,\tilde\Omega_+,\tilde\Omega_-$ can be tuned independently by changing the amplitudes of the driving lasers; some particular Hamiltonians for specific values of these amplitudes are shown in Table \ref{tab:drives}.
-Third, due to the appearance of mutually commuting pairs of Hamiltonians in Table \ref{tab:drives}, specifically $-\s_\alpha^2$ and $\pm \s_\alpha+\s_\alpha^2$ for $\alpha\in\set{\z,\x}$, the three-laser drive admits pulse sequences that exactly implement arbitrary SU($2$) (spatial) rotations of the form $e^{-\i\chi\vec n\cdot\vec s}$, where $\chi$ is a rotation angle, $\vec n$ is a rotation axis, and $\vec s\equiv(\s_\x,\s_\y,\s_\z)$.
+Third, due to the appearance of mutually commuting pairs of Hamiltonians in Table \ref{tab:drives}, specifically $-\s_\alpha^2$ and $\pm \s_\alpha+\s_\alpha^2$ for $\alpha\in\set{\z,\x}$, the three-laser drive admits pulse sequences that exactly implement arbitrary SU(2) (spatial) rotations of the form $e^{-\i\chi\vec n\cdot\vec s}$, where $\chi$ is a rotation angle, $\vec n$ is a rotation axis, and $\vec s\equiv(\s_\x,\s_\y,\s_\z)$.
 The capability to perform arbitrary spatial rotations, together with the capability to measure the number of atoms with spin projection $\mu$ onto a fixed quantization axis, $\bk{\S_{\mu\mu}}$ (where $\S_{\mu\nu}=\sum_j\s_{\mu\nu j}$), implies the capability to reconstruct all components of the mean collective spin matrix $\bk{\v\S}=\sum_{\mu\nu}\bk{\S_{\mu\nu}}\op{\mu}{\nu}$ via spin qudit tomography \cite{newton1968measurability, perlin2020qudit}.
-Moreover, we expect that advanced quantum optimal control techniques (similar to those of Refs.~\cite{anderson2015accurate, lucarelli2018quantum}) can be used to implement arbitrary SU($n$) rotations by designing suitable time-dependent drive amplitudes.
+Moreover, we expect that advanced quantum control techniques (similar to those of Refs.~\cite{anderson2015accurate, lucarelli2018quantum}) can be used to implement arbitrary SU($n$) rotations by designing suitable time-dependent drive amplitudes.
 
@@
 \end{align}
-For two-level spins with $\mu=\pm\frac12$, $\hat w_{+,q}^{(\phi)}$ is proportional to the identity operator and $\hat w_{+,q}^{(\phi)} = 2\sin\p{\phi/2} \s_{\z,q}$, so the kinetic Hamiltonian in the gauge frame describes a (synthetic) inhomogeneous magnetic field:
+For two-level spins with $\mu=\pm\frac12$, $\hat w_{+,q}^{(\phi)}$ is proportional to the identity operator and $\hat w_{-,q}^{(\phi)} = 2\sin\p{\phi/2} \s_{\z,q}$, so the kinetic Hamiltonian in the gauge frame describes a (synthetic) inhomogeneous magnetic field:
 \begin{align}
@@
 Finally, the gauge transformation $\c_{q\mu}^\dag\to \c_{q-\mu\phi,\mu}^\dag$ also transforms the interaction Hamiltonian.
-Applying this transformation to Eq.~\eqref{eq:H_int_momenta} and keeping only terms that respect the coherences imposed on initial states by the laser drive, we can again derive an effective spin model.
+Applying this transformation to Eq.~\eqref{eq:H_int_momenta} and keeping only terms that respect coherences that can be imposed on initial states by the laser drive in Section \ref{sec:controls} (applied to an initially spin-down-polarized state) again results in an effective spin model.
 For sufficiently weak SOC ($s\phi\to0$) this spin model is still well-approximated by $\H_{\t{int}}$ in Eqs.~\eqref{eq:H_int_start} and \eqref{eq:H_int}.
@@
 The weak-field effective Hamiltonian preserves permutational symmetry, so $\bbk{\sds}_\MF\to1$ as $h\to0$.
-Moreover, the initial y-magnetization $\bk{\S_\y}=0$ is conserved by $\H_\spin$, so long-time-averaged magnetization $\sigma_\MF$ is determined by the time-average of $\s_\x$ for a single spin:
+Moreover, the initial y-magnetization $\bk{\S_\y}=0$ is conserved by $\H_\spin$, so the long-time-averaged magnetization $\sigma_\MF$ is determined by the time-average of $\s_\x$ for a single (any) spin:
 \begin{align}
@@
 Given the possibility for long-range SU($n$) interactions, we hope our work stimulates further efforts into simulating SY and SYK-like models \cite{sachdev1993gapless, bentsen2019integrable} in cold atomic platforms.
-In follow-up work, it would be interesting to study the the relationship between initial states and dynamical phases of our SU($n$) spin model more systematically, and to consider the effect of quantum corrections to mean-field behavior.
+In follow-up work, it would be interesting to study the relationship between initial states and dynamical phases of our SU($n$) spin model more systematically, and to consider the effect of quantum corrections to mean-field behavior.
 There is also room to improve on the three-laser drive introduced in this work, for which it is natural to ask what additional techniques or ingredients are necessary to implement universal control of individual nuclear spins.
@@
 
-We thank V.~Gurarie and E.~Yuzbashyan for useful discussions on the topic.
-This work was supported by the AFOSR grants FA9550-18-1-0319, FA9550-19-1-027, by the DARPA and ARO grant W911NF-16-1-0576, ARO W911NF-19-1-0210, DOE (QSA), NSF PHY1820885, NSF JILA-PFC PHY-1734006 and NSF QLCI-2016244 grants, and by NIST.
+We thank Victor Gurarie, Emil Yuzbashyan, Asier P.~Orioli, and Jeremy T.~Young for helpful discussions on this work.
+This work was supported by the AFOSR grants FA9550-18-1-0319, FA9550-19-1-027, by the DARPA and ARO grant W911NF-16-1-0576, ARO W911NF-19-1-0210, DOE (QSA), NSF PHY1820885, NSF JILA-PFC PHY-1734006, NSF QLCI-2016244, and by NIST.
 

### 2021-12-16 sun_phases/qudit_tomo/qudit_tomo.tex
moving qudit tomo materials

--- 2021-09-30 sun_phases/qudit_tomo.tex
@@
 \usepackage{graphicx} % for figures
-\graphicspath{{./figures/qudit_tomo/}} % set path for all figures
+\graphicspath{{./figures/}} % set path for all figures
 
@@
 After introducing a basis of operators closely related to the spherical harmonics, which obey the rotational symmetries of spin qudits, we map our quantum tomography task onto the classical problem of signal recovery on the sphere.
-We then provide algorithms with $O\p{rd^3}$ serial runtime, parallelizable down to $O\p{rd^2}$, for (i) computing a priori upper bounds on the expected error with which spin projection measurements along $r$ given axes can reconstruct an unknown qudit state, and (ii) estimating a posteriori the statistical error in a reconstructed state.
+We then provide algorithms with $O(rd^3)$ serial runtime, parallelizable down to $O(rd^2)$, for (i) computing a priori upper bounds on the expected error with which spin projection measurements along $r$ given axes can reconstruct an unknown qudit state, and (ii) estimating a posteriori the statistical error in a reconstructed state.
 Our algorithms motivate a simple randomized tomography protocol, for which we find that using more measurement axes can yield substantial benefits that plateau after $r\approx3d$.
@@
 Due to advancements in experimental capabilities to address nuclear spin states (i.e.~hyperfine levels) in ultracold atomic systems \cite{daley2011quantum, lu2011strongly, mischuck2012control, aikawa2012boseeinstein, smith2013quantum, cazalilla2014ultracold}, as well as developments in the control of ultracold molecular systems \cite{bohn2017cold, takekoshi2014ultracold, zeppenfeld2012sisyphus, kozyryev2017sisyphus, puri2017synthesis, wu2017cryofuge, marco2019degenerate, liu2019molecular, anderegg2019optical, chou2020frequencycomb, lin2020quantum}, a particular setting of growing interest is the spin qudit, or a multilevel quantum angular momentum degree of freedom.
-Spin qudits can provide advantages over their qubit counterparts for quantum sensing \cite{hemmer2018squeezing, evrard2019enhanced}, enable quantum simulations of SU($d$) magnetism \cite{cazalilla2014ultracold, banerjee2013atomic, zhang2014spectroscopic, scazza2014observation, goban2018emergence, perlin2021engineering}, and offer unique capabilities for quantum computation and error correction \cite{albert2020robust, gross2021designing, barnes2021assembly}.
+Spin qudits can provide advantages over their qubit counterparts for quantum sensing \cite{evrard2019enhanced, hemmer2021squeezing}, enable quantum simulations of SU($d$) magnetism \cite{cazalilla2014ultracold, banerjee2013atomic, zhang2014spectroscopic, scazza2014observation, goban2018emergence, perlin2021engineering}, and offer unique capabilities for quantum computation and error correction \cite{albert2020robust, gross2021designing, barnes2021assembly}.
 In all cases, quantum state tomography is necessary to take full advantage of a spin qudit\footnote{Note that the measurement of collective observables for quantum sensing or simulation can be recast as a single-spin tomography task.}.
@@
 Moreover, the degree-$\ell$ polarization operators $T_{\ell m}$ transform similarly to spin-$\ell$ particles and spherical harmonics $Y_{\ell m}$ under 3D rotations (see Appendix \ref{sec:rotations}).
-Specifically, for any triplet of angles $\bm\omega=\p{\alpha,\beta,\gamma}$, we can therefore define the rotation operator
+Specifically, for any triplet of angles $\bm\omega=\p{\alpha,\beta,\gamma}$, we can define the rotation operator
 \begin{align}
@@
 which is a stronger version of the condition that we found for the spin qudit tomography problem in Eq.~\eqref{eq:tomo_recovery}.
-We will refer to Eq.~\eqref{eq:full_recovery} as the {\it full recovery problem}, and Eq.~\eqref{eq:tomo_recovery} as the the {\it reduced recovery problem}.
+We will refer to Eq.~\eqref{eq:full_recovery} as the {\it full recovery problem}, and Eq.~\eqref{eq:tomo_recovery} as the {\it reduced recovery problem}.
 Due to the fact that $D^\ell_{0,m}\p{\bm v} = \sqrt{\frac{4\pi}{2\ell+1}}\, Y_{\ell m}\p{\bm v}$, any solution to the full recovery problem automatically solves the reduced recovery problem by setting $C_{\ell m}\p{\bm v} = \sqrt{\frac{2\ell+1}{4\pi}}\, \tilde C_{\ell m}\p{\bm v}^*$.
@@
 The measurement matrix $M_V$ allows us to make concrete statements about the statistical error between the empirical estimate $\tilde\rho_V$ and the true state $\rho$.
-Assume, for example, that the estimates $\tilde T_j$ are equal to $\bk{T_j}_\rho$ up to uncorrelated noise with variance no grater than $\epsilon^2$:
+Assume, for example, that the estimates $\tilde T_j$ are equal to $\bk{T_j}_\rho$ up to uncorrelated noise with variance no greater than $\epsilon^2$:
 \begin{align}
@@
 \item For each axis $\bm v\in V$, make $n$ measurements of spin projection, and set $\tilde\Pi_{\bm v\mu}\approx\bk{\Pi_{\bm v\mu}}_\rho$ to the fraction of times in which the measurement outcome was $\mu$.
-\item Use the the estimates $\tilde\Pi_{\bm v\mu}$ of $\bk{\Pi_{\bm v\mu}}_\rho$ to compute estimates of $\bk{T_{\bm v\ell,0}}_\rho$,
+\item Use the estimates $\tilde\Pi_{\bm v\mu}$ of $\bk{\Pi_{\bm v\mu}}_\rho$ to compute estimates of $\bk{T_{\bm v\ell,0}}_\rho$,
 \begin{align}
@@
   \caption{Empirical measurement-adjusted error scales $\tilde\beta(p)$ with $p$ excess measurement axes, determined by minimizing over $10^3$ choices of measurement axes $V$ or 5 minutes of runtime (for each $p$), whichever comes first.
-    Color indicates the qudit dimension $d$.
+    Marker and color indicates the qudit dimension $d$.
     The rapid initial drop in $\tilde\beta(p)$ implies that using more measurement axes can substantially lower the upper bound on reconstruction error provided in Eq.~\eqref{eq:bound}, and that these benefits plateau after $p\approx d$.
@@
 The method of Ref.~\cite{newton1968measurability} therefore requires extremely careful fine-tuning of measurement axis orientations for large spin dimensions.
-For this reason, we expect our randomized tomography protocol be be more robust to errors in axis orientation.
+For this reason, we expect our randomized tomography protocol to be more robust to errors in axis orientation.
 We leave a detailed analysis of robustness to errors in axis orientation and the effect of these errors on state reconstruction to future work.
@@
 
-Denoting the state of a spin-$s$ particle spin spin projection $\mu$ onto a quantization axis by $\ket{s\mu}$, we define
+Denoting the state of a spin-$s$ particle with spin projection $\mu$ onto a quantization axis by $\ket{s\mu}$, we define
 \begin{align}

### 2022-10-04 sun_phases/spin_model_paper/main.tex
updating spin model paper materials

@@
 \newcommand{\Bk}{\Braket} % shorthand for braket notation
-\renewcommand{\v}{\bm} % bold vectors
 \renewcommand{\d}{\partial} % partial derivative
@@
 \newcommand{\bbk}[1]{\langle\!\langle #1 \rangle\!\rangle}
-\newcommand{\Bbk}[1]
-{\left\langle\!\!\left\langle #1 \right\rangle\!\!\right\rangle}
+\newcommand{\Bbk}[1]{\left\langle\!\!\left\langle #1 \right\rangle\!\!\right\rangle}
 
@@
 % symbols with hats
-\renewcommand{\b}{\hat b}
-\renewcommand{\c}{\hat c}
-\newcommand{\s}{\hat s}
-\renewcommand{\H}{\hat H}
-\renewcommand{\S}{\hat S}
-\newcommand{\R}{\hat R}
-\newcommand{\T}{\hat T}
+\renewcommand{\b}{\hat{b}}
+\renewcommand{\c}{\hat{c}}
+\newcommand{\s}{\hat{s}}
+\renewcommand{\H}{\hat{H}}
+\renewcommand{\S}{\hat{S}}
+\newcommand{\R}{\hat{R}}
+\newcommand{\T}{\hat{T}}
 \renewcommand{\O}{\hat{\mathcal{O}}}
@@
 
-\newcommand{\xx}{\x\x}
-\newcommand{\xxi}{\x\x_\i}
+\newcommand{\xx}{{\x\x}}
+\newcommand{\xxi}{{\x\x_\i}}
 \newcommand{\X}{\text{X}}
-\newcommand{\XX}{\X\X}
-\newcommand{\XXI}{\X\X_\i}
+\newcommand{\XX}{{\X\X}}
+\newcommand{\XXI}{{\X\X_\i}}
 \newcommand{\spin}{\text{spin}}
@@
 \newcommand{\su}{\mathfrak{su}}
-\newcommand{\sds}{\bar{\v s}\cdot\bar{\v s}}
+\newcommand{\sds}{\bar{\bm{s}}\cdot\bar{\bm{s}}}
 \DeclareMathOperator{\diag}{diag}
@@
 \graphicspath{{./figures/}} % set path for all figures
+\usepackage[caption=false]{subfig} % subfigures (via \subfloat[]{})
+\newcommand{\sref}[1]{\protect\subref{#1}} % for referencing subfigures
 
@@
 
-\date{22 September 2021}
+\date{15 December 2021}
 
@@
 Underpinning much of high energy physics, the SU($n$) gauge theory known as Yang-Mills theory is central to our understanding of the electroweak and strong forces.
-Extensions of Yang-Mills and SU($n$) symmetry feature in the most well-studied examples of holographic duality \cite{maldacena1999largen} and the connection between entanglement and gravity \cite{ryu2006holographic} through the anti-de Sitter/conformal field theory (AdS/CFT) correspondence.
+Extensions of Yang-Mills and SU($n$) symmetry feature in the most well-studied examples of holographic duality \cite{maldacena1999largen} and the connection between entanglement and gravity \cite{ryu2006holographic} through the anti-de Sitter / conformal field theory (AdS/CFT) correspondence.
 In a condensed matter setting, SU(2) appears ubiquitously as a symmetry of the Hubbard model, with important consequences for the study of quantum magnetism and high temperature superconductivity \cite{lee2006doping}.
-The extension of SU(2) Hubbard and spin models to SU($n$) has led to predictions of exotic phases of matter such as valence bond solids \cite{read1989valencebond, rokhsar1990quadratic, kaul2012lattice, hermele2011topological} and chiral spin liquids \cite{hermele2009mott, hermele2011topological, chen2016syntheticgaugefield, nataf2016chiral}, as well as the potential to perform universal topological quantum computation \cite{freedman2004class, nayak2008nonabelian} and other phenomena \cite{nataf2014exact, nataf2016exact}.
+The extension of SU(2) Hubbard and spin models to SU($n$) has led to predictions of exotic phases of matter such as valence bond solids \cite{read1989valencebond, rokhsar1990quadratic, kaul2012lattice, hermele2011topological} and chiral spin liquids \cite{hermele2009mott, hermele2011topological, chen2016syntheticgaugefield, nataf2016chiral}, novel magnetic behaviors \cite{nataf2014exact, nataf2016exact}, and flavor selective phase transitions \cite{xu2018interaction, delre2018selective, tusi2021flavourselective, ibarra-garcia-padilla2021universal} as well as the potential to perform universal topological quantum computation \cite{freedman2004class, nayak2008nonabelian}.
 Furthermore, disordered SU($n$) spin models have opened analytically tractable avenues for studying quantum chaos and information scrambling \cite{sachdev1993gapless}.
@@
 As a result, AMO experiments can directly probe the role of SU($n$) interactions in controllable settings.
-Recent progress includes studies of the thermodynamic properties of SU($n$) fermionic gases \cite{hazzard2012hightemperature, bonnes2012adiabatic, stellmer2013degenerate, yip2014theory, pagano2014onedimensional, choudhury2020collective, song2020evidence, sonderhouse2020thermodynamics}, SU($n$) Hubbard phases and phase transitions \cite{taie2012su, hofrichter2016direct, taie2020observation}, single- \cite{messio2012entropy} and two-orbital \cite{cappellini2014direct, scazza2014observation, zhang2014spectroscopic, beverland2016realizing} SU($n$) magnetism, and multi-body SU($n$)-symmetric interactions \cite{goban2018emergence, perlin2019effective}.
+Recent progress includes studies of the thermodynamic properties of SU($n$) fermionic gases \cite{hazzard2012hightemperature, bonnes2012adiabatic, stellmer2013degenerate, yip2014theory, ibarra-garcia-padilla2021universal, pagano2014onedimensional, choudhury2020collective, song2020evidence, sonderhouse2020thermodynamics}, SU($n$) Hubbard phases and phase transitions \cite{taie2012su, hofrichter2016direct, taie2020observation}, single- \cite{messio2012entropy} and two-orbital SU($n$) magnetism \cite{cappellini2014direct, scazza2014observation, zhang2014spectroscopic, beverland2016realizing}, and multi-body SU($n$)-symmetric interactions \cite{goban2018emergence, perlin2019effective}.
 
@@
 \end{align}
-where $N$ is the total number of atoms on the lattice, we define $u\equiv U\times N/L$ for convenience, $\delta_{k+p,\ell+q}=1$ if $k+p=\ell+q$ and zero otherwise (enforcing conservation of momentum).
-
-If the interaction energy $U$ is small compared to the single-particle bandwidth $4J$, then the mode-changing collisions in $\H_{\t{int}}$ become off-resonant, motivating the frozen-mode approximation $\set{k,p}=\set{\ell,q}$ (i.e.~either $k=\ell$ and $p=q$, or $k=q$ and $p=\ell$)\footnote{Note that the frozen-mode approximation neglects correlated momentum-hopping terms of the form $\c_{\pi-p,\mu}^\dag \c_{\pi-q,\mu} \c_{p\nu}^\dag \c_{q\nu}$, which conserve both momentum and energy.
-We defer a careful treatment of these terms to future work, noting only that they vanish on the manifold of permutationally symmetric spin states with one atom per lattice site, and that the frozen-mode approximation is benchmarked in Refs.~\cite{he2019engineering, smale2019observation} and Appendix \ref{sec:benchmarking}.}.
+where $N$ is the total number of atoms on the lattice, we define $u\equiv U\times N/L$ for convenience, and the Kronecker delta $\delta_{k+p,\ell+q}=1$ if $k+p=\ell+q$ and zero otherwise (enforcing conservation of momentum).
+
+Momentum-space modes are delocalized across the lattice, which is why interactions between these modes are $O(U/L)=O(u/N)$.
+Mode-changing collisions thereby generally become off-resonant when these interactions are weak compared to the single-particle mode spacing, i.e.~$U/L\ll J/L$ (equivalently $U\ll J$)\footnote{
+A reasonable objection to this argument is that each momentum-space atom interacts with $O(N)$ other atoms, which suggests that its total interaction energy is $O(U)\gg J/L$.
+However, the \textit{typical} energy difference between two states that are coupled by interactions is $O(J)$, so if $U\ll J$ then again interactions become off-resonant in the ``typical'' case, leaving only the ``atypical'' terms that approximately conserve energy, which again leads to the frozen-mode approximation.
+}.
+In this regime, we can make the frozen-mode approximation $\set{k,p}=\set{\ell,q}$ (i.e.~either $k=\ell$ and $p=q$, or $k=q$ and $p=\ell$)\footnote{
+Note that the frozen-mode approximation neglects correlated momentum-hopping terms of the form $\c_{\pi-p,\mu}^\dag \c_{\pi-q,\mu} \c_{p\nu}^\dag \c_{q\nu}$, which also conserve energy.
+We defer a careful treatment of these terms to future work, noting only (i) that they vanish on the manifold of permutationally symmetric spin states with one atom per lattice site, and (ii) that the frozen-mode approximation is benchmarked in Refs.~\cite{he2019engineering, smale2019observation} and Appendix \ref{sec:benchmarking}.
+}.
 The terms with $k=\ell$ and $p=q$ are $\frac{u}{2N} \sum \c_{k\mu}^\dag \c_{k\mu} \c_{p\nu}^\dag \c_{p\nu} = \frac12 N u$, which is a constant  energy shift that we can freely neglect.
@@
 \begin{align}
-  \v\s_q \equiv \sum_{\mu,\nu} \s_{\mu\nu q} \op{\mu}{\nu},
-\end{align}
-and for any pair of such operator-valued matrices $\v{\hat{A}},\v{\hat{B}}$, we define the inner product
-\begin{align}
-  \v{\hat{A}} \cdot \v{\hat{B}}
-  \equiv \sum_{\mu,\nu} \hat A_{\mu\nu}^\dag \hat B_{\mu\nu}.
+  \bm\s_q \equiv \sum_{\mu,\nu} \s_{\mu\nu q} \op{\mu}{\nu},
+\end{align}
+and for any pair of such operator-valued matrices $\bm{\hat{A}},\bm{\hat{B}}$, we define the inner product
+\begin{align}
+  \bm{\hat{A}} \cdot \bm{\hat{B}}
+  \equiv \sum_{\mu,\nu} \hat{A}_{\mu\nu}^\dag \hat{B}_{\mu\nu}.
   \label{eq:dot_product}
@@
 \begin{align}
-  \H_{\t{int}} = -\f{u}{2N} \sum_{p,q} \v\s_p\cdot\v\s_q
-  = -\f{u}{2N}\v\S\cdot\v\S,
+  \H_{\t{int}} = -\f{u}{2N} \sum_{p,q} \bm\s_p\cdot\bm\s_q
+  = -\f{u}{2N}\bm\S\cdot\bm\S,
   \label{eq:H_int}
 \end{align}
-where $\v\S\equiv\sum_q\v\s_q$ is a collective spin matrix, analogous to the collective spin vector $\vec S=(\S_\x,\S_\y,\S_\z)$ in the case of SU(2) \cite{he2019engineering}, with $\frac12\v\S\cdot\v\S \simeq \vec S\cdot\vec S = \S_\x^2 + \S_\y^2 + \S_\z^2$ when $n=2$ (here $\simeq$ denotes equality up to identity terms).
+where $\bm\S\equiv\sum_q\bm\s_q$ is a collective spin matrix, analogous to the collective spin vector $\vec S=(\S_\x,\S_\y,\S_\z)$ in the case of SU(2) \cite{he2019engineering}, with $\frac12\bm\S\cdot\bm\S \simeq \vec S\cdot\vec S = \S_\x^2 + \S_\y^2 + \S_\z^2$ when $n=2$ (here $\simeq$ denotes equality up to identity terms).
 
@@
 (a) Ultracold atoms on a lattice of $L$ sites tunnel between neighboring lattice sites at a rate $J$, and locally repel each other with interaction energy $U$.
-(b,c) When the interaction energy $U$ is small compared to the single-particle bandwidth $4J$, the frozen-mode approximation enables the interaction Hamiltonian to be written as a spin model consisting of exchange terms $\v\s_p\cdot\v\s_q$, which swap the states of two spins pinned to modes $p,q$.
+(b,c) When the interaction energy $U$ is small compared to the single-particle bandwidth $4J$, the frozen-mode approximation enables the interaction Hamiltonian to be written as a spin model consisting of exchange terms $\bm\s_p\cdot\bm\s_q$, which swap the states of two spins pinned to modes $p,q$.
 (d) Interactions open an energy gap $u=U\times N/L$ between the manifold of permutationally symmetric states of $N$ spins, and the orthogonal complement of states that break spin-permutation symmetry.
@@
 We now discuss the spin Hamiltonian $\H_{\t{int}}$ in Eq.~\eqref{eq:H_int}.
-The operator $\v\s_p\cdot\v\s_q$ simply swaps the nuclear spin states of two atoms pinned to modes $p,q$.
-The term $-\v\s_p\cdot\v\s_q$ thereby assigns a definite energy of $-1$ ($+1$) to a pair of spins that are symmetric (anti-symmetric) under exchange.
-In this sense, $\v\s_p\cdot\v\s_q$ is analogous to the enforcement of SU(2) spin alignment by ferromagnetic interactions, which similarly assigns different energies to the anti-symmetric spin-0 singlet $\ket{\up\dn}-\ket{\dn\up}$ and the symmetric spin-1 triplets $\set{\ket{\up\up},\ket{\dn\dn},\ket{\up\dn}+\ket{\dn\up}}$.
-By summing over all pair-wise exchange terms $\v\s_p\cdot\v\s_q$, the interaction Hamiltonian $\H_{\t{int}}$ energetically enforces a permutational symmetry among all spins, opening an energy gap $u$ between the manifold of all permutationally symmetric (PS) states and the orthogonal complement of excited (e.g.~spin-wave) states that break permutational symmetry.
+The operator $\bm\s_p\cdot\bm\s_q$ simply swaps the nuclear spin states of two atoms pinned to modes $p,q$.
+The term $-\bm\s_p\cdot\bm\s_q$ thereby assigns a definite energy of $-1$ ($+1$) to a pair of spins that are symmetric (anti-symmetric) under exchange.
+In this sense, $\bm\s_p\cdot\bm\s_q$ is analogous to the enforcement of SU(2) spin alignment by ferromagnetic interactions, which similarly assigns different energies to the anti-symmetric spin-0 singlet $\ket{\up\dn}-\ket{\dn\up}$ and the symmetric spin-1 triplets $\set{\ket{\up\up},\ket{\dn\dn},\ket{\up\dn}+\ket{\dn\up}}$.
+By summing over all pair-wise exchange terms $\bm\s_p\cdot\bm\s_q$, the interaction Hamiltonian $\H_{\t{int}}$ energetically enforces a permutational symmetry among all spins, opening an energy gap $u$ between the manifold of all permutationally symmetric (PS) states and the orthogonal complement of excited (e.g.~spin-wave) states that break permutational symmetry.
 See Figure \ref{fig:spin_model} for a summary of this section thus far.
@@
 More generally, the state $\ket\psi$ of an $n$-level spin can be represented by a quasi-probability distribution $Q_\psi$ on the Bloch sphere (commonly known as the Husimi-$Q$ function, e.g.~in the spin-squeezing community \cite{ma2011quantum}).
-The value $Q_\psi\p{\v v}$ at a point $\v v$ on the sphere is equal to the overlap of $\ket\psi$ with a pure state $\ket{\v v}$ that is maximally polarized in the direction of $\v v$: $Q_\psi\p{\v v}\equiv\abs{\bk{\v v|\psi}}^2$ (see Figure \ref{fig:spin_dist}).
-In the case of a mixed state $\hat\rho$, this distribution is defined by $Q_{\hat\rho}\p{\v v}\equiv\bk{\v v|\hat\rho|\v v}$.
+The value $Q_\psi\p{\bm{v}}$ at a point $\bm{v}$ on the sphere is equal to the overlap of $\ket\psi$ with a pure state $\ket{\bm{v}}$ that is maximally polarized in the direction of $\bm{v}$: $Q_\psi\p{\bm{v}}\equiv\abs{\bk{\bm{v}|\psi}}^2$ (see Figure \ref{fig:spin_dist}).
+In the case of a mixed state $\hat\rho$, this distribution is defined by $Q_{\hat\rho}\p{\bm{v}}\equiv\bk{\bm{v}|\hat\rho|\bm{v}}$.
 Closely related spherical representations of multilevel spin states and operators are discussed in Refs.~\cite{dowling1994wigner, li2013weylwignermoyal}.
@@
 We consider a specific three-laser driving scheme with a geometry sketched in Figure \ref{fig:3LD}.
-Here the lattice lies in the $y$-$z$ plane at an angle $\theta$ to the $z$ axis, oriented along $\v\ell=(0,\sin\theta,\cos\theta)$.
+Here the lattice lies in the $y$-$z$ plane at an angle $\theta$ to the $z$ axis, oriented along $\bm\ell=(0,\sin\theta,\cos\theta)$.
 We set the spin quantization axis along $z$.
@@
 \begin{enumerate*}
-\item two counter-propagating right-circularly polarized lasers with drive amplitudes $\Omega_\pm$ and wavevectors $\kappa\v v_\pm$, propagating in opposite directions along the $z$ axis, $\v v_\pm=(0,0,\pm1)$, and
-\item a third laser linearly polarized along $z$, with drive amplitude $\Omega_0$ and wavevector $\kappa \v v_0$, propagating along the $x$ axis $\v v_0=(1,0,0)$.
+\item two counter-propagating right-circularly polarized lasers with drive amplitudes $\Omega_\pm$ and wavevectors $\kappa\bm{v}_\pm$, propagating in opposite directions along the $z$ axis, $\bm{v}_\pm=(0,0,\pm1)$, and
+\item a third laser linearly polarized along $z$, with drive amplitude $\Omega_0$ and wavevector $\kappa \bm{v}_0$, propagating along the $x$ axis $\bm{v}_0=(1,0,0)$.
 \end{enumerate*}
@@
   \p{e^{-\i m\phi j} \s_{mj} \otimes\op{\e}{\g}_j + \t{h.c.}}
-  + \Delta \hat N_\e,
-\end{align}
-where $m\in\set{+1,0,-1}$ indexes the laser pointing along $\v v_m$; the SOC angle $\phi\equiv\kappa\v v_+\cdot\v\ell=\kappa \cos\theta$ (in units with lattice spacing $a=1$); $\s_{\z,j},\s_{+,j}\s_{-,j}$ are standard axial, spin-raising, and spin-lowering operators for the spin at lattice site $j$; $\s_{0,j}\equiv\s_{\z,j}$ for shorthand; $\ket\g_j$ and $\ket\e_j$ respectively denote the ground and excited electronic states of atom $j$; and $\hat N_\e=\1\otimes\sum_j\op\e_j$ counts the number of excited atoms (with $\1$ the identity operator on all spin degrees of freedom).
+  + \Delta \hat{N}_\e,
+\end{align}
+where $m\in\set{+1,0,-1}$ indexes the laser pointing along $\bm{v}_m$; the SOC angle $\phi\equiv\kappa\bm{v}_+\cdot\bm\ell=\kappa \cos\theta$ (in units with lattice spacing $a=1$); $\s_{\z,j},\s_{+,j}\s_{-,j}$ are standard axial, spin-raising, and spin-lowering operators for the spin at lattice site $j$; $\s_{0,j}\equiv\s_{\z,j}$ for shorthand; $\ket\g_j$ and $\ket\e_j$ respectively denote the ground and excited electronic states of atom $j$; and $\hat{N}_\e=\1\otimes\sum_j\op\e_j$ counts the number of excited atoms (with $\1$ the identity operator on all spin degrees of freedom).
 
@@
 Third, due to the appearance of mutually commuting pairs of Hamiltonians in Table \ref{tab:drives}, specifically $-\s_\alpha^2$ and $\pm \s_\alpha+\s_\alpha^2$ for $\alpha\in\set{\z,\x}$, the three-laser drive admits pulse sequences that exactly implement arbitrary SU(2) (spatial) rotations of the form $e^{-\i\chi\vec n\cdot\vec s}$, where $\chi$ is a rotation angle, $\vec n$ is a rotation axis, and $\vec s\equiv(\s_\x,\s_\y,\s_\z)$.
-The capability to perform arbitrary spatial rotations, together with the capability to measure the number of atoms with spin projection $\mu$ onto a fixed quantization axis, $\bk{\S_{\mu\mu}}$ (where $\S_{\mu\nu}=\sum_j\s_{\mu\nu j}$), implies the capability to reconstruct all components of the mean collective spin matrix $\bk{\v\S}=\sum_{\mu\nu}\bk{\S_{\mu\nu}}\op{\mu}{\nu}$ via spin qudit tomography \cite{newton1968measurability, perlin2020qudit}.
+The capability to perform arbitrary spatial rotations, together with the capability to measure the number of atoms with spin projection $\mu$ onto a fixed quantization axis, $\bk{\S_{\mu\mu}}$ (where $\S_{\mu\nu}=\sum_j\s_{\mu\nu j}$), implies the capability to reconstruct all components of the mean collective spin matrix $\bk{\bm\S}=\sum_{\mu\nu}\bk{\S_{\mu\nu}}\op{\mu}{\nu}$ via spin qudit tomography \cite{newton1968measurability, perlin2021spin}.
 Moreover, we expect that advanced quantum control techniques (similar to those of Refs.~\cite{anderson2015accurate, lucarelli2018quantum}) can be used to implement arbitrary SU($n$) rotations by designing suitable time-dependent drive amplitudes.
@@
 We assume the capability to prepare an initial state in which all spins are maximally polarized along the $z$ axis, i.e.~$\ket{\z}^{\otimes N}=\ket{s}^{\otimes N}$, which is unaffected by the gauge transformation (up to a global phase).
-The three-laser then allows us to rotate this state into one that is polarized along any spatial axis (in the gauge frame).
+The best-reported fidelity for preparing a spin-polarized degenerate Fermi gas is about $\sim92\%$ \cite{sonderhouse2020thermodynamics}, and there are no major technical barriers to further improvement.
+After preparing a spin-polarized gas, the three-laser drive then allows us to rotate this state into one that is polarized along any spatial axis (in the gauge frame).
 In addition, when $n>2$ the three-laser drive allows us to prepare product states with nontrivial intra-spin correlations.
@@
 Before discussing SOC for $n$-level fermions, we briefly review the well-studied case of two-level SOC with a one-dimensional lattice \cite{wall2016synthetic, kolkowitz2016spinorbitcoupled, bromley2018dynamics, he2019engineering}.
-In this case, SOC is induced by an external driving field that imprints a phase $e^{-\i\phi j}$ on lattice site $j$, or equivalently imparts a momentum kick $q\to q+\phi$, upon the absorption of a photon\footnote{In order for the drive Hamiltonian $\H_{\t{drive}}^{(\phi)}$ to be well-defined, $\phi$ should be commensurate with the lattice, e.g.~$\phi\in\ZZ\times2\pi/L$ on a one-dimensional lattice of $L$ sites.}:
+In this case, SOC is induced by an external driving field that imprints a phase $e^{-\i\phi j}$ on lattice site $j$, or equivalently imparts a momentum kick $q\to q+\phi$, upon the absorption of a photon\footnote{
+In order for the drive Hamiltonian $\H_{\t{drive}}^{(\phi)}$ to be well-defined, $\phi$ should be commensurate with the lattice, e.g.~$\phi\in\ZZ\times2\pi/L$ on a one-dimensional lattice of $L$ sites.
+}:
 \begin{align}
@@
 In contrast, the drive we considered in Section \ref{sec:controls} addresses electronic excitations off-resonantly, inducing an effective Hamiltonian in the ground-state hyperfine manifold with spin projections $\mu\in\set{s,s-1,\cdots,-s}$ (a similar scheme was used to study SOC in a subspace of the ground-state manifold in Ref.~\cite{mancini2015observation}).
-Nonetheless, both the two-level drive in Eq.~\eqref{eq:drive_2} and the $n$-level drive in Eq.~\eqref{eq:H_3LD} become homogeneous (i.e.~independent of the spatial mode index $j$ or $q$) and independent of the SOC angle $\phi$ after the same spin-symmetric gauge transformation\footnote{The ``asymmetric'' gauge transformation $(\c_{j,\up}^\dag,\c_{j,\dn}^\dag)\to(e^{\i\phi j} \c_{j,\up}^\dag,\c_{j,\dn}^\dag)$, sometimes performed in the two-state SOC literature, does not generalize as nicely to $n>2$.} $\c_{j\mu}^\dag\to e^{\i\phi\mu j} \c_{j\mu}^\dag$.
+Nonetheless, both the two-level drive in Eq.~\eqref{eq:drive_2} and the $n$-level drive in Eq.~\eqref{eq:H_3LD} become homogeneous (i.e.~independent of the spatial mode index $j$ or $q$) and independent of the SOC angle $\phi$ after the same spin-symmetric gauge transformation\footnote{
+The ``asymmetric'' gauge transformation $(\c_{j,\up}^\dag,\c_{j,\dn}^\dag)\to(e^{\i\phi j} \c_{j,\up}^\dag,\c_{j,\dn}^\dag)$, sometimes performed in the two-state SOC literature, does not generalize as nicely to $n>2$.
+} $\c_{j\mu}^\dag\to e^{\i\phi\mu j} \c_{j\mu}^\dag$.
 
@@
   = -2J \sum_q
-  \sp{\cos\p{q} \hat w_{+,q}^{(\phi)} - \sin\p{q} \hat w_{-,q}^{(\phi)}},
+  \sp{\cos\p{q} \hat{w}_{+,q}^{(\phi)} - \sin\p{q} \hat{w}_{-,q}^{(\phi)}},
 \end{align}
@@
 \begin{align}
-  \hat w_{+,q}^{(\phi)} &\equiv \sum_\mu \cos\p{\mu\phi} \s_{\mu\mu q}, \\
-  \hat w_{-,q}^{(\phi)} &\equiv \sum_\mu \sin\p{\mu\phi} \s_{\mu\mu q}.
-\end{align}
-For two-level spins with $\mu=\pm\frac12$, $\hat w_{+,q}^{(\phi)}$ is proportional to the identity operator and $\hat w_{-,q}^{(\phi)} = 2\sin\p{\phi/2} \s_{\z,q}$, so the kinetic Hamiltonian in the gauge frame describes a (synthetic) inhomogeneous magnetic field:
+  \hat{w}_{+,q}^{(\phi)} &\equiv \sum_\mu \cos\p{\mu\phi} \s_{\mu\mu q}, \\
+  \hat{w}_{-,q}^{(\phi)} &\equiv \sum_\mu \sin\p{\mu\phi} \s_{\mu\mu q}.
+\end{align}
+For two-level spins with $\mu=\pm\frac12$, $\hat{w}_{+,q}^{(\phi)}$ is proportional to the identity operator and $\hat{w}_{-,q}^{(\phi)} = 2\sin\p{\phi/2} \s_{\z,q}$, so the kinetic Hamiltonian in the gauge frame describes a (synthetic) inhomogeneous magnetic field:
 \begin{align}
@@
 \begin{align}
-  \H_\spin = -\f{u}{2N} \v\S\cdot\v\S + 2J\phi \sum_q \sin\p{q} \s_{\z,q},
+  \H_\spin = -\f{u}{2N} \bm\S\cdot\bm\S + 2J\phi \sum_q \sin\p{q} \s_{\z,q},
   \label{eq:H_spin}
 \end{align}
-consisting of a spin-locking $\v\S\cdot\v\S$ term that energetically favors permutational symmetry, and an inhomogeneous magnetic field that causes inter-spin dephasing.
+consisting of a spin-locking $\bm\S\cdot\bm\S$ term that energetically favors permutational symmetry, and an inhomogeneous magnetic field that causes inter-spin dephasing.
 
@@
 We now study the dynamical behavior of the SOC spin Hamiltonian $H_\spin$ in Eq.~\eqref{eq:H_spin}, and henceforth work exclusively in the ``gauge frame'' of $\H_\spin$ and the three-laser drive $\H_{\t{3LD}}$ in Eq.~\eqref{eq:H_3LD}.
-We use a Ramsey-like setup wherein we prepare an initial state with the three-laser drive (using fast pule sequences), then let the state evolve freely for some time under $\H_\spin$, and finally apply again the three-laser drive to map observables of interest onto spin projection measurements (e.g.~with spin qudit tomography \cite{newton1968measurability, perlin2020qudit}).
+We use a Ramsey-like setup wherein we prepare an initial state with the three-laser drive (using fast pulse sequences), then let the state evolve freely for some time under $\H_\spin$, and finally apply again the three-laser drive to map observables of interest onto spin projection measurements (e.g.~with spin qudit tomography \cite{newton1968measurability, perlin2021spin}).
 At the mean-field (MF) level, the undriven spin Hamiltonian (neglecting constant energy shifts) becomes
 \begin{align}
-  \H_\MF = u \sum_q\sp{-\bk{\bar{\v s}}\cdot\v\s_q
+  \H_\MF = u \sum_q\sp{-\bk{\bar{\bm{s}}}\cdot\bm\s_q
     + h \sin\p{q} \s_{\z,q}}.
@@
 \end{align}
-where $\bar{\v s}\equiv\frac1N\sum_q\v\s_q$ is the average spin matrix, and $h\equiv 2J\phi/u$ is a dimensionless strength of the inhomogeneous magnetic field.
+where $\bar{\bm{s}}\equiv\frac1N\sum_q\bm\s_q$ is the average spin matrix, and $h\equiv 2J\phi/u$ is a dimensionless strength of the inhomogeneous magnetic field.
 We assume that all momenta $q\in\ZZ_N\times 2\pi/N$ are occupied.
@@
 \begin{align}
-  \bbk{\sds}_\MF = \f1{N^2} \times \bbk{\v\S\cdot\v\S}_\MF.
+  \bbk{\sds}_\MF = \f1{N^2} \times \bbk{\bm\S\cdot\bm\S}_\MF.
 \end{align}
 By design, these non-negative quantities are normalized to lie on the interval $[0,1]$, independent of the system size $N$ or spin dimension $n$.
-In the remainder of this section we will assume that $n$ is even, both for the sake of experimental relevance (most relevant atomic nuclei are fermionic) and to avoid complications from parity effects.
+In the remainder of this section we will assume that $n$ is even, both for the sake of experimental relevance (most relevant atomic nuclei are fermionic) and to avoid complications from parity effects\footnote{
+Odd $n$ is accessible in principle by addressing a subset of the hyperfine levels of an even-$n$ atom.
+However, the controls in Section \ref{sec:controls} unavoidably address \textit{all} hyperfine sublevels of an atom.
+These controls are used to prepare the initial states in Section \ref{sec:mean_field}, and are moreover the source of SOC in Section \ref{sec:SOC} (which is in turn the origin of the spin Hamiltonian simulated in Section \ref{sec:mean_field}).
+Preparing analogous initial states and simulating analogous spin Hamiltonians to those in Section \ref{sec:mean_field} is therefore nontrivial quantum state and Hamiltonian engineering problem.
+}.
 
@@
 
-\begin{figure}
+\begin{figure*}
 \centering
-\includegraphics{mean_mag_int_X.pdf}
+\includegraphics{mean_mag_int.pdf}
 \caption{
-Time-averaged magnetization $\sigma_\MF$ and interaction energy $\bbk{\sds}_\MF$ for different spin dimensions $n$ (indicated in the legend) as determined by mean-field simulations of $N=100$ spins initially in the x-polarized state $\ket\X$ for a time $T=10^5/u$.
-Insets show same data after rescaling $h\to h \times (n/2)^{1/3}$, and transforming vertical axes according to Eq.~\eqref{eq:rescale}.
+Time-averaged magnetization $\sigma_\MF$ (top row) and interaction energy $\bbk{\sds}_\MF$ (bottom row) for initial states $\ket\X$ (left column), $\ket\XX$ (middle column) and $\ket\XXI$ (right column) and different spin dimensions $n$ (indicated in the top legend),
+as determined by mean-field simulations of $N=100$ spins for a time $T=10^5/u$.
+The initial states are $N$-fold tensor products of the states defined in Eqs.~\eqref{eq:state_x} and \eqref{eq:states_xx_xxi}.
+Insets show the same data after rescaling $h\to h \times (n/2)^{1/3}$, and transforming vertical axes according to Eq.~\eqref{eq:rescale} (with the exception of the $\bbk{\sds}_\MF$ for the initial state $\ket\XX$, in the bottom row of the middle column, which is shifted using $2\gamma(n)$ rather than $\gamma(n)$; inset data for $\bbk{\sds}_\MF$ with $\ket\XX$ and $n=2$ is excluded altogether due to a division by $0$).
+All insets have the same axis limits.
 }
-\label{fig:mean_mag_int_X}
-\end{figure}
-
-Figure \ref{fig:mean_mag_int_X} shows the time-averages of the magnetization $\sigma_\MF$ and interaction energy $\bbk{\sds}_\MF$ as computed by mean-field simulations of $N=100$ spins initially in the x-polarized state $\ket\X\equiv\ket\x^{\otimes N}$, where
+\label{fig:mean_mag_int}
+\end{figure*}
+
+Figure \ref{fig:mean_mag_int} shows time-averages of the mean-field magnetization $\sigma_\MF$ and interaction energy $\bbk{\sds}_\MF$ for $N=100$ spins in a few different initial states.
+We first discuss the case of an initial x-polarized state $\ket\X\equiv\ket\x^{\otimes N}$, where
 \begin{align}
@@
 The ordered phase has a non-zero magnetization $\sigma_\MF$ and an interaction energy $\bbk{\sds}_\MF$ that asymptotically approach their maximal values as $h\to0$.
-The disordered phase has no (time-averaged) magnetization, $\sigma_\MF=0$, but the interaction energy $\bbk{\sds}_\MF$ nonetheless indicates persistent nontrivial inter-spin correlations when $n>2$.
-These nontrivial correlations vanish as $h\to\infty$, in which case $\bbk{\sds}_\MF$ approaches the minimal value allowed by conservation laws (clarified below).
+The disordered phase has no (time-averaged) magnetization, $\sigma_\MF=0$, but the interaction energy $\bbk{\sds}_\MF$ nonetheless indicates persistent inter-spin correlations that tend to a minimal value enforced by conservation laws as $h\to\infty$ (clarified below).
 By minimizing the reduced field $h$ for which $\sigma_\MF=0$, we numerically find that the transition between ordered and disordered phases occurs at a critical field $h_\crit=\p{n/2}^{-\alpha}$ with $\alpha\approx1/3$ (see Figure \ref{fig:crit_fields_X}).
@@
 
-As shown in insets of Figure \ref{fig:mean_mag_int_X}, mean-field results for different spin dimensions $n$ collapse onto each other when normalizing the field $h$ to its critical value, $h\to h\times\p{n/2}^{1/3}$, and rescaling
+As shown in the insets of the left column in Figure \ref{fig:mean_mag_int}, mean-field results for different spin dimensions $n$ collapse onto each other when normalizing the field $h$ to its critical value, $h\to h\times\p{n/2}^{1/3}$, and rescaling
 \begin{align}
@@
 The time-averaged transverse magnetization $\sigma_\MF$ then trivially vanishes as $h\to\infty$.
-The interaction energy $\bk{\sds}_\MF = \bk{\bar{\v s}}_\MF \cdot \bk{\bar{\v s}}_\MF + O(1/N)$, meanwhile, has contributions from:
+The interaction energy $\bk{\sds}_\MF = \bk{\bar{\bm{s}}}_\MF \cdot \bk{\bar{\bm{s}}}_\MF + O(1/N)$, meanwhile, has contributions from:
 \begin{enumerate*}
-\item the diagonal parts of the mean spin matrix $\bk{\bar{\v s}}_\MF$, which are conserved by inhomogeneous spin precession, and
-\item the off-diagonal parts of $\bk{\bar{\v s}}_\MF$, whose oscillations average to zero when evaluating the time average in $\bbk{\sds}_\MF$.
+\item the diagonal parts of the mean spin matrix $\bk{\bar{\bm{s}}}_\MF$, which are conserved by inhomogeneous spin precession, and
+\item the off-diagonal parts of $\bk{\bar{\bm{s}}}_\MF$, whose oscillations average to zero when evaluating the time average in $\bbk{\sds}_\MF$.
 \end{enumerate*}
-Altogether, the interaction energy $\bbk{\sds}_\MF$ in the strong-field limit is determined by the time-independent diagonal part $\diag\bk{\bar{\v s}}_\MF = \diag\op{\x}$, namely
+Altogether, the interaction energy $\bbk{\sds}_\MF$ in the strong-field limit is determined by the time-independent diagonal part $\diag\bk{\bar{\bm{s}}}_\MF = \diag\op{\x}$, namely
 \begin{align}
@@
 
-In the weak-field limit $h\to0$, the spin-locking $\v\S\cdot\v\S$ interactions of the Hamiltonian $\H_\spin$ energetically restrict dynamics to the permutationally symmetric (PS) manifold.
+In the weak-field limit $h\to0$, the spin-locking $\bm\S\cdot\bm\S$ interactions of the Hamiltonian $\H_\spin$ energetically restrict dynamics to the permutationally symmetric (PS) manifold.
 To first order in $h$, the effect of the inhomogeneous field can be acquired by projecting it onto the PS manifold, which takes $\s_{\z,q}\to\frac1N\S_\z$.
@@
 \end{align}
-We can adapt exact analytical results for the dynamics of an infinite-range Ising model \cite{foss-feig2013nonequilibrium}\footnote{See Appendix K of Ref.~\cite{perlin2020shorttime} for a simpler adaptation of the analytics in Ref.~\cite{foss-feig2013nonequilibrium} to the one-axis twisting model $\H_{\t{OAT}}=\chi \s_\z^2$.} to find that
+We can adapt exact analytical results for the dynamics of an infinite-range Ising model \cite{foss-feig2013nonequilibrium}\footnote{
+See Appendix K of Ref.~\cite{perlin2020shorttime} for a simpler adaptation of the analytics in Ref.~\cite{foss-feig2013nonequilibrium} to the one-axis twisting model $\H_{\t{OAT}}=\chi \s_\z^2$.
+} to find that
 \begin{align}
@@
 
+Finally, we note that the mean-field interaction energy $\bbk{\sds}_\MF$ exhibits a sharp transition for $n=2$ that appears to get smoothed out as $n$ increases.
+This behavior likely has to do with the fact that mean-field theory is sensitive to additional intra-spin correlations and coherences when $n>2$, and thereby sensitive to more quantum fluctuations.
+Quantum fluctuations are known to smooth out signatures of a dynamical phase transition, and moreover these fluctuations have a larger effect on higher-order correlators \cite{lerose2019impact}.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\begin{figure}
-\centering
-\includegraphics{mean_mag_int_XX.pdf}
-\caption{
-A corollary of Figure \ref{fig:mean_mag_int_X} for the initial state $\ket\XX$.
-The inset for interaction energy $\bbk{\sds}_\MF$ in Figure \ref{fig:mean_mag_int_X} subtracts off the minimal value of $\bbk{\sds}_\MF$ and rescales to lie on the interval $[0,1]$, as prescribed in Eq.~\eqref{eq:rescale}.
-Here the subtracting and rescaling is identical, but with a minimal value of $2\gamma\p{n}$ rather than $\gamma\p{n}$.
-}
-\label{fig:mean_mag_int_XX}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{mean_mag_int_XXI.pdf}
-\caption{
-A corollary of Figure \ref{fig:mean_mag_int_X} for the initial state $\ket\XXI$.
-Insets show the same data shifted and rescaled identically to Figure \ref{fig:mean_mag_int_X}.
-}
-\label{fig:mean_mag_int_XXI}
-\end{figure}
-
-We now consider the same setup as above, but with the initial ``kitten'' states $\ket\XX\equiv\ket\xx^{\otimes N}$ and $\ket\XXI\equiv\ket\xxi^{\otimes N}$, where
+We now discuss the results in the middle and right columns of Figure \ref{fig:mean_mag_int}, for the initial ``kitten'' states $\ket\XX\equiv\ket\xx^{\otimes N}$ and $\ket\XXI\equiv\ket\xxi^{\otimes N}$, where
 \begin{align}
@@
   \ket\xxi \equiv \f{\ket{\x} + \p{-1}^s \ket{-\x}}{\sqrt{2}},
+  \label{eq:states_xx_xxi}
 \end{align}
@@
 \end{align}
-Similarly to Figure \ref{fig:mean_mag_int_X}, Figures \ref{fig:mean_mag_int_XX} and \ref{fig:mean_mag_int_XXI} show the time-averaged magnetization $\sigma_\MF$ and interaction energy \mbox{$\bbk{\sds}_\MF$} throughout mean-field dynamics of the initial states $\ket\XX$ and $\ket\XXI$.
-These figures exclude the trivial case of spin dimension $n=2$, for which $\ket\xx=\ket{-s}$ is an eigenstate of $\H_\spin$ and $\ket\xxi=e^{-\i\frac{\pi}{2}\s_\z}\ket\x\equiv\ket\y$ is spin-polarized along the $y$ axis.
-The first and perhaps most interesting observation to make about Figures \ref{fig:mean_mag_int_XX} and \ref{fig:mean_mag_int_XXI} is that they are different, signifying the importance of intra-spin coherences for the dynamical behavior of multilevel spin models.
-
-Unlike Figure \ref{fig:mean_mag_int_X} (for $\ket\X$), Figure \ref{fig:mean_mag_int_XX} (for $\ket\XX$) exhibits no sharp transition between distinct dynamical phases: the time-averaged magnetization $\sigma_\MF=0$ for all values of the field $h$, and the interaction energy $\bbk{\sds}_\MF$ smoothly crosses over from a maximal value of 1 to a minimal value of $2\gamma\p{n}$.
+Here $s$ and $\mu$ are half-integer-valued, so $(-1)^s=\pm\i$ and $(-1)^{s+\mu}=\pm1$.
+The first and perhaps most interesting observation to make about the results for $\ket\XX$ and $\ket\XXI$ in the middle and right columns of Figure \ref{fig:mean_mag_int} is that they are different, signifying the importance of intra-spin coherences for the dynamical behavior of multilevel spin models.
+
+Unlike the case of $\ket\X$, the results for $\ket\XX$ exhibit no sharp transition between distinct dynamical phases.
+When $n=2$, the kitten state $\ket\XX$ is polarized along $-z$, and is therefore an eigenstate of $\H_\spin$ with $\sigma_\MF=\bbk{\sds}_\MF=1$ at all times.
+When $n>2$, the time-averaged magnetization $\sigma_\MF=0$ for all values of the field $h$, and the interaction energy $\bbk{\sds}_\MF$ smoothly crosses over from a maximal value of 1 to a minimal value of $2\gamma\p{n}$.
 The minimal value of $\bbk{\sds}_\MF$ approached as $h\to\infty$ can be explained with arguments identical to those in the paragraph containing Eq.~\eqref{eq:lim_ss}, which now imply that
@@
   = 2\gamma\p{n}.
-\end{align}
-The vanishing magnetization $\sigma_\MF=0$ in Figure \ref{fig:mean_mag_int_XX} is protected by symmetries of $\H_\spin$ and $\ket\XX$.
-For all initial states that we have considered, the value of $\bk{\S_\z}=0$ is conserved by the spin Hamiltonian $\H_\spin$.
+  \label{eq:lim_ss_XX}
+\end{align}
+The vanishing initial magnetization $\sigma_\MF=0$ for $\ket\XX$ with $n>2$, meanwhile, is protected by symmetries of $\H_\spin$ and $\ket\XX$.
+The collective spin operator $\S_\z$ commutes with the spin Hamiltonian $\H_\spin$, so $\bk{S_\z}=0$ at all times for all initial states we consider (with the exception of $\ket\XX$ with $n=2$, for which $\bk{\S_\z}=-\frac{N}{2}$).
 Moreover, both the spin Hamiltonian $\H_\spin$ and the state $\ket\XX$ are invariant (up to global phase) under the action of $\R_\z^\pi$, where $\R_\z^\theta\equiv e^{-\i\theta \S_\z}$, which is to say that
@@
 where $\simeq$ denotes equality up to an overall phase.
-This symmetry implies that
+This symmetry implies that, for the initial state $\ket\XX$,
 \begin{align}
@@
 
-Turning now to mean-field results for the initial kitten state $\ket\XXI$ in Figure \ref{fig:mean_mag_int_XXI}, we remark that the magnetization $\sigma_\MF$ and interaction energy $\bbk{\sds}_\MF$ behave identically to those for the initial spin-polarized state $\ket\X$ in Figure \ref{fig:mean_mag_int_X}.
+Turning now to mean-field results for the initial kitten state $\ket\XXI$, we remark that the magnetization $\sigma_\MF$ and interaction energy $\bbk{\sds}_\MF$ behave identically to those for the initial spin-polarized state $\ket\X$.
 This finding can be understood through the fact that
@@
 
-To make sense of why the magnetization $\sigma_\MF$ is identical in Figure \ref{fig:mean_mag_int_XXI} for $\ket\XXI$ as in Figure \ref{fig:mean_mag_int_X} for $\ket\X$, we follow a four-part argument:
+To make sense of why the magnetization $\sigma_\MF$ is identical for an initial state $\ket\XXI$ as for $\ket\X$, we follow a four-part argument:
 \begin{enumerate}
-\item \label{pt:spin_mat} The time-averaged magnetization vector $\bbk{\vec\sigma}_\MF$ can be written as a function of the time-averaged spin matrix $\bbk{\bar{\v s}}_\MF$.
-\item \label{pt:diags} The spin matrix $\bbk{\bar{\v s}}_\MF$ is only ever nonzero on its diagonal and anti-diagonal, regardless of the initial state.
-That is, nonzero components $\bbk{\bar s_{\mu\nu}}_\MF$ of $\bbk{\bar{\v s}}_\MF$ always have $\mu=\pm\nu$ (see discussion below).
-\item \label{pt:twist} The twist operator $\T_\z^\theta$ acts trivially on the diagonal and anti-diagonal components of $\bar{\v s}$, which together with point \ref{pt:diags} implies that $\bbk{\T_\z^\theta{}^\dag \, \bar{\v s} \, \T_\z^\theta}_\MF = \bbk{\bar{\v s}}_\MF$.
+\item \label{pt:spin_mat} The time-averaged magnetization vector $\bbk{\vec\sigma}_\MF$ can be written as a function of the time-averaged spin matrix $\bbk{\bar{\bm{s}}}_\MF$.
+\item \label{pt:diags} The spin matrix $\bbk{\bar{\bm{s}}}_\MF$ is only ever nonzero on its diagonal and anti-diagonal, regardless of the initial state.
+That is, nonzero components $\bbk{\bar s_{\mu\nu}}_\MF$ of $\bbk{\bar{\bm{s}}}_\MF$ always have $\mu=\pm\nu$ (see discussion below).
+\item \label{pt:twist} The twist operator $\T_\z^\theta$ acts trivially on the diagonal and anti-diagonal components of $\bar{\bm{s}}$, which together with point \ref{pt:diags} implies that $\bbk{\T_\z^\theta{}^\dag \, \bar{\bm{s}} \, \T_\z^\theta}_\MF = \bbk{\bar{\bm{s}}}_\MF$.
 \item \label{pt:rot} The rotation operator $\R_\z^\theta$ merely rotates the magnetization vector $\bbk{\vec\sigma}_\MF$ without changing its magnitude.
@@
 The only nontrivial step in the above argument is point \ref{pt:diags}, which says that $\bbk{\bar s_{\mu\nu}}_\MF$ is guaranteed to be zero unless $\mu=\pm\nu$.
-This observation, nominally a numerical result of mean-field simulations, can be understood as follows.
-The eigenstates $\ket{m,w}$ of $\H_\spin$ are uniquely identified by definite numbers $m=(m_s,m_{s-1},\cdots,m_{-s})$ of atoms occupying each internal spin state $\mu\in\set{s,s-1,\cdots,-s}$, and an auxiliary index $w$ that encodes how $\ket{m,w}$ transforms under permutations of all spins (see Appendix \ref{sec:pert_theory})\footnote{Seen otherwise, since $\S_{\mu\mu}$ commutes with $\H_\spin$, eigenvectors of $\H_\spin$ can be indexed by eigenvalues of $\S_{\mu\mu}$.
-The number $m_\mu$ is then the eigenvalue of $\ket{m,w}$ with respect to $\S_{\mu\mu}$, i.e.~$\S_{\mu\mu}\ket{m,w} = m_\mu\ket{m,w}$, while $w$ encodes all other information required to uniquely specify $\ket{m,w}$.}.
+This observation, nominally a numerical finding in mean-field simulations, can be understood as follows.
+The eigenstates $\ket{m,w}$ of $\H_\spin$ are uniquely identified by definite numbers $m=(m_s,m_{s-1},\cdots,m_{-s})$ of atoms occupying each internal spin state $\mu\in\set{s,s-1,\cdots,-s}$, and an auxiliary index $w$ that encodes how $\ket{m,w}$ transforms under permutations of all spins (see Appendix \ref{sec:pert_theory})\footnote{
+Seen otherwise, since $\S_{\mu\mu}$ commutes with $\H_\spin$, eigenvectors of $\H_\spin$ can be indexed by eigenvalues of $\S_{\mu\mu}$.
+The number $m_\mu$ is then the eigenvalue of $\ket{m,w}$ with respect to $\S_{\mu\mu}$, i.e.~$\S_{\mu\mu}\ket{m,w} = m_\mu\ket{m,w}$, while $w$ encodes all other information required to uniquely specify $\ket{m,w}$.
+}.
 The operator $\bar s_{\mu\nu}=\frac1N \S_{\mu\nu}$ with $\mu\ne\nu$ couples the state $\ket{m,w}$ to states $\ket{m',w'}$ in which $(m_\mu',m'_\nu)=(m_\mu+1,m_\nu-1)$.
@@
 In the weak-field limit $h\to0$, such a degeneracy occurs at the mean-field level between PS states differing only in the populations $m_\mu,m_{-\mu}$ (with a fixed value of $m_\mu+m_{-\mu}$), as the effective Hamiltonian becomes $\H_\MF^\eff\propto\sum_\mu\mu^2m_\mu$.
-This symmetry is preserved at all orders in perturbation theory\footnote{Only even powers of the ``perturbation'' $\sum_q\sin\p{q}\s_{\z,q}$ can be nonzero within the PS manifold, and even powers of this perturbation exhibit the same mean-field degeneracy between states differing only in the populations $m_\mu,m_{-\mu}$.}, so some coherence between such states is preserved as $h\to h_\crit$, although this coherence decays as perturbative corrections to degenerate eigenstates cause them to leak out of the PS manifold (and thereby have a smaller overlap with the initial state $\ket\X$).
-Note that beyond-mean-field effects break the symmetry protecting anti-diagonal components of $\bbk{\v\s}_\MF$, causing them to decay on time scales that should diverge as $N\to\infty$.
+This symmetry is preserved at all orders in perturbation theory\footnote{
+Only even powers of the ``perturbation'' $\sum_q\sin\p{q}\s_{\z,q}$ can be nonzero within the PS manifold, and even powers of this perturbation exhibit the same mean-field degeneracy between states differing only in the populations $m_\mu,m_{-\mu}$.
+}, so some coherence between such states is preserved as $h\to h_\crit$, although this coherence decays as perturbative corrections to degenerate eigenstates cause them to leak out of the PS manifold (and thereby have a smaller overlap with the initial state $\ket\X$).
+Note that beyond-mean-field effects break the symmetry protecting anti-diagonal components of $\bbk{\bm\s}_\MF$, causing them to decay on time scales that should diverge as $N\to\infty$.
+
+As a final point, we note that the numerical results in this work are obtained at zero temperature.
+A nonzero motional temperature $T\lesssim\Delta_{\t{gap}}$, where $\Delta_{\t{gap}}$ is the single-particle band gap, should have little effect on our results: at a filling of $f=N/L=1$ atoms per lattice site, the initial state and Hamiltonian are essentially identical for $T=0$ and $T\lesssim\Delta_{\t{gap}}$ (with exponentially small corrections).
+A filling $f<1$, meanwhile, a non-zero temperature results in a mixture of different occupied quasi-momenta, which corresponds to single-particle $\s_\z$ disorder for the spin model in Eq.~\eqref{eq:H_spin}\footnote{
+Note, however, that $f\approx1$ is necessary for the validity of the spin model at any temperature.
+}.
+This disorder does not effect the phenomenology of the spin model, which is insensitive to the precise form of the inhomogeneous single-particle field.
+Finally, a nonzero spin temperature should result in a spin mixture that suppresses the magnitudes of $\bbk{\vec\sigma}$ and $\bbk{\sds}$.
+We leave a more detailed analysis of nonzero temperatures to future work.
 
@@
 Our work makes important progress in understanding the SU($n$) Fermi-Hubbard model in experimentally relevant parameter regimes, and we expect our findings to be readily testable in experiments with ultracold atoms.
-Given the possibility for long-range SU($n$) interactions, we hope our work stimulates further efforts into simulating SY and SYK-like models \cite{sachdev1993gapless, bentsen2019integrable} in cold atomic platforms.
+Given the possibility for long-range SU($n$) interactions, we hope our work stimulates further efforts into simulating Sachdev-Ye-like (SY-like) and Sachdev-Ye-Kitaev-like (SYK-like) models \cite{sachdev1993gapless, bentsen2019integrable} in cold atomic platforms.
 In follow-up work, it would be interesting to study the relationship between initial states and dynamical phases of our SU($n$) spin model more systematically, and to consider the effect of quantum corrections to mean-field behavior.
@@
 Universal control would allow for an experimental study of $n$-dependence (including even/odd-$n$ parity effects) in a single experimental platform, simply by controlling the occupation and coherence of internal spin states.
-Finally, one can also study the SU($n$) Hubbard model in the super-exchange regime that gives rise to a real-space (as opposed to momentum-space) spin model, where SOC gives rise to chiral multilevel spin interactions.
-Unlike our present work, the super-exchange regime does not require weak SOC, and therefore has a larger parameter space in which to explore dynamical behavior.
+Finally, one can also study the SU($n$) Hubbard model in the superexchange regime that gives rise to a real-space (as opposed to momentum-space) spin model, where SOC gives rise to chiral multilevel spin interactions.
+Unlike our present work, the superexchange regime does not require weak SOC, and therefore has a larger parameter space in which to explore dynamical behavior.
 
@@
 We thank Victor Gurarie, Emil Yuzbashyan, Asier P.~Orioli, and Jeremy T.~Young for helpful discussions on this work.
-This work was supported by the AFOSR grants FA9550-18-1-0319, FA9550-19-1-027, by the DARPA and ARO grant W911NF-16-1-0576, ARO W911NF-19-1-0210, DOE (QSA), NSF PHY1820885, NSF JILA-PFC PHY-1734006, NSF QLCI-2016244, and by NIST.
+This work was supported by the AFOSR grants FA9550-18-1-0319, FA9550-19-1-0275, by the DARPA and ARO grant W911NF-16-1-0576, ARO W911NF-19-1-0210, NSF PHY1820885, NSF JILA-PFC PHY-1734006, QLCI Award OMA - 2016244, the U.S.
+Department of Energy, Office of Science, National Quantum Information Science Research Centers, Quantum Systems Accelerator (QSA) and by the National Institute of Standards and Technology (NIST).
 
@@
 In this appendix we present numerical evidence to support the validity of the spin models derived in Sections \ref{sec:spin_model} and \ref{sec:SOC}.
-Figures \ref{fig:benchmarking_SU4} and \ref{fig:benchmarking_SU6} show a set of time-averaged observables computed with numerically exact simulations of a Fermi-Hubbard model and an effective spin model, respectively, with $n=4$ (Figure \ref{fig:benchmarking_SU4}) and $n=6$ (Figure \ref{fig:benchmarking_SU6}) internal levels per spin.
+Figures \ref{fig:benchmarking_SU4} and \ref{fig:benchmarking_SU6} show a set of time-averaged observables computed via numerical integration of the Schr\"odinger equation for a Fermi-Hubbard model and an effective spin model, respectively, with $n=4$ (Figure \ref{fig:benchmarking_SU4}) and $n=6$ (Figure \ref{fig:benchmarking_SU6}) internal levels per spin.
 Details for these simulations are provided in the caption of Figure \ref{fig:benchmarking_SU4}.
 Our main conclusion from these figures is that the two models show remarkable agreement for the observables considered in our work.
-Note that these results are only intended to benchmark the approximation of a Fermi-Hubbard model by a spin model; these results are not expected to agree with the mean-field theory in Section \ref{sec:mean_field} due to strong finite-size effects.
-
-\begin{figure}[!h]
+Due to strong finite-size effects, the results in Figures \ref{fig:benchmarking_SU4} and \ref{fig:benchmarking_SU6} do not provide reliable numerical values for realistic systems, and are only intended to benchmark the approximation of a Fermi-Hubbard model by a spin model.
+In particular, these results are not expected to agree with the mean-field theory in Section \ref{sec:mean_field}.
+
+\begin{figure}[h]
 \centering
@@
 \caption{
-Numerical results (analogous to Figures \ref{fig:mean_mag_int_X}, \ref{fig:mean_mag_int_XX}, and \ref{fig:mean_mag_int_XXI} of the main text) for the time-averaged interaction energy and magnetization (both normalized to a maximal value of 1) in a system of $L=5$ lattice sites, for both a Fermi-Hubbard model (dots) and spin model (lines) with $n=4$ internal states per spin.
+Numerical results (analogous to Figure \ref{fig:mean_mag_int} of the main text) for the time-averaged interaction energy and magnetization (both normalized to a maximal value of 1) in a system of $L=5$ lattice sites, for both a Fermi-Hubbard model (dots) and spin model (lines) with $n=4$ internal states per spin.
 The corresponding initial state (defined in Section \ref{sec:mean_field} of the main text) is indicated in each panel, and observables are averaged over a time $tJ=200$.
@@
 Results for the initial kitten state $\ket\XXI$ are excluded because they are identical to those of $\ket\X$, and magnetization for the initial state $\ket\XX$ is always $0$.
-Note that while panels (a) and (b) are representative of infinite-time behavior, the inset in panel (c) shows that the Fermi-Hubbard and spin models exhibit different behaviors on very long time scales, although good agreement is restored by rescaling time in the spin model, indicating the likelihood of a need to renormalize spin model parameters.
+Note that while panels (a) and (b) are representative of infinite-time behavior, the inset in panel (c) shows that the spin model (dotted line) and Fermi-Hubbard model (solid line instead of dots in the inset) exhibit different behaviors on very long time scales, although good agreement is restored by rescaling time in the spin model, indicating the likelihood of a need to renormalize spin model parameters.
 In any case, such time scales are inaccessible in current experiments and diverge as $N\to\infty$, so these corrections do not affect the main results of our work.
@@
 
-\begin{figure}[!h]
+\begin{figure}[h]
 \centering
@@
   &&
-  \hat\Pi_{ij} \equiv \v\s_i\cdot\v\s_j
+  \hat\Pi_{ij} \equiv \bm\s_i\cdot\bm\s_j
   = \sum_{\mu,\nu} \s_{\mu\nu i} \s_{\nu\mu j},
@@
 \end{align}
-or more generally an $M$-body operator\footnote{At face value, an $M$-body operator with $M>2$ does not typically appear in experiments.
-Nonetheless, considering $M>2$ illuminates the structure of eigenstates (and eigenvalues) of $\H_0$, and allows us to go to high orders in perturbation theory with single- and two-body perturbations.}
-\begin{align}
-  \O(w,\hat X) = \sum_{k\in\D_N\p{M}} w_k \hat X_k,
+or more generally an $M$-body operator\footnote{
+At face value, an $M$-body operator with $M>2$ does not typically appear in experiments.
+Nonetheless, considering $M>2$ illuminates the structure of eigenstates (and eigenvalues) of $\H_0$, and allows us to go to high orders in perturbation theory with single- and two-body perturbations.
+}
+\begin{align}
+  \O(w,\hat{X}) = \sum_{k\in\D_N\p{M}} w_k \hat{X}_k,
 \end{align}
@@
 with $M=2$; $k\equiv\p{k_1,k_2,\cdots,k_M}$ is a list of the
-individual spins $k_i\in\ZZ_N\equiv\set{1,2,\cdots,N}$ that the operator $\hat X_k\equiv \hat X_{k_1k_2\cdots k_M}$ acts on; and
+individual spins $k_i\in\ZZ_N\equiv\set{1,2,\cdots,N}$ that the operator $\hat{X}_k\equiv \hat{X}_{k_1k_2\cdots k_M}$ acts on; and
 \begin{align}
@@
 \end{align}
-is the strictly ``off-diagonal'' part of $\ZZ_N^M$, which is necessary to identify for a consistent definition of $\hat X_k$ as an $M$-body operator.
-In this notation, the magnetic field and Ising Hamiltonians in Eq.~\eqref{eq:field_ising} respectively become $\O\p{B,\s_{\z}}$ and $\O\p{J,\s_\z\otimes \s_\z}$.
-
-If the addition $\O(w,\hat X)$ to the SU($n$)-symmetric Hamiltonian $\H_0$ in Eq.~\eqref{eq:H_0} is sufficiently small, namely with operator norm $\norm*{\O(w,\hat X)}$ less than half the spectral gap $\Delta_{\t{gap}}$ of $\H_0$, $\norm*{\O(w,\hat X)}<\Delta_{\t{gap}}/2$, then we can treat the effect of $\O(w,\hat X)$ on the ground-state PS manifold $\E_0$ perturbatively.
-The effective Hamiltonians $\H_\eff^{(1)}$ and $\H_\eff^{(2)}$ induced by $\O(w,\hat X)$ on the PS manifold $\E_0$ at leading orders in perturbation theory are \cite{bravyi2011schrieffer}
-\begin{align}
-  \H_\eff^{(1)} = \P_0 \O(w,\hat X) \P_0,
+is the strictly ``off-diagonal'' part of $\ZZ_N^M$, which is necessary to identify for a consistent definition of $\hat{X}_k$ as an $M$-body operator.
+In this notation, the magnetic field and Ising Hamiltonians in Eq.~\eqref{eq:field_ising} respectively become $\O\p{B,\s_\z}$ and $\O\p{J,\s_\z\otimes \s_\z}$.
+
+If the addition $\O(w,\hat{X})$ to the SU($n$)-symmetric Hamiltonian $\H_0$ in Eq.~\eqref{eq:H_0} is sufficiently small, namely with operator norm $\norm*{\O(w,\hat{X})}$ less than half the spectral gap $\Delta_{\t{gap}}$ of $\H_0$, $\norm*{\O(w,\hat{X})}<\Delta_{\t{gap}}/2$, then we can treat the effect of $\O(w,\hat{X})$ on the ground-state PS manifold $\E_0$ perturbatively.
+The effective Hamiltonians $\H_\eff^{(1)}$ and $\H_\eff^{(2)}$ induced by $\O(w,\hat{X})$ on the PS manifold $\E_0$ at leading orders in perturbation theory are \cite{bravyi2011schrieffer}
+\begin{align}
+  \H_\eff^{(1)} = \P_0 \O(w,\hat{X}) \P_0,
   &&
   \H_\eff^{(2)} = - \sum_{\Delta\ne0}
-  \f1\Delta \P_0 \O(w,\hat X) \P_\Delta \O(w,\hat X) \P_0,
+  \f1\Delta \P_0 \O(w,\hat{X}) \P_\Delta \O(w,\hat{X}) \P_0,
   \label{eq:H_eff}
@@
 where $\P_\Delta$ is a projector onto the eigenspace $\E_\Delta$ of $\H_0$ with interaction energy $\Delta$ above that of the PS manifold.
-The first order effective Hamiltonian $\H_\eff^{(1)}$ simply projects $\O(w,\hat X)$ onto the PS manifold $\E_0$, and takes the form
+The first order effective Hamiltonian $\H_\eff^{(1)}$ simply projects $\O(w,\hat{X})$ onto the PS manifold $\E_0$, and takes the form
 \begin{align}
@@
   &&
-  \col X \equiv \sum_{k\in\D_N\p{M}} \hat X_k,
+  \col X \equiv \sum_{k\in\D_N\p{M}} \hat{X}_k,
   \label{eq:mean_col}
@@
 The second order effective Hamiltonian $\H_\eff^{(2)}$ in Eq.~\eqref{eq:H_eff} takes more work to simplify due to the presence of a projector $\P_\Delta$ onto the manifold $\E_\Delta$ of states with excitation energy $\Delta$.
-This projector essentially picks off the part of $\O(w,\hat X)$ that is strictly off-diagonal with respect to the ground- and excited-state manifolds $\E_0$ and $\E_\Delta$.
-We therefore need to decompose $\O(w,\hat X)$ into components that generate states of definite excitation energy when acting on PS states $\ket\psi\in\E_0$.
+This projector essentially picks off the part of $\O(w,\hat{X})$ that is strictly off-diagonal with respect to the ground- and excited-state manifolds $\E_0$ and $\E_\Delta$.
+We therefore need to decompose $\O(w,\hat{X})$ into components that generate states of definite excitation energy when acting on PS states $\ket\psi\in\E_0$.
 The SU($n$) symmetry of $\H_0$ enables such a decomposition to take the form
 \begin{align}
-  \H_0 \O(w,\hat X) \ket\psi
-  = \sum_\Delta \p{E_0+\Delta} \O(w^\Delta,\hat X) \ket\psi,
+  \H_0 \O(w,\hat{X}) \ket\psi
+  = \sum_\Delta \p{E_0+\Delta} \O(w^\Delta,\hat{X}) \ket\psi,
   &&
@@
 
-Equipped with the decomposition $\O(w,\hat X)=\sum_\Delta\O(w^\Delta,\hat X)$ with terms $\O(w^\Delta,\hat X)$ that generate states of definite excitation energy $\Delta$, we can expand
+Equipped with the decomposition $\O(w,\hat{X})=\sum_\Delta\O(w^\Delta,\hat{X})$ with terms $\O(w^\Delta,\hat{X})$ that generate states of definite excitation energy $\Delta$, we can expand
 \begin{align}
   \H_{\t{eff}}^{(2)} = -\sum_{\Delta\ne0} \f1\Delta
-  \P_0 \O(w^\Delta,\hat X)^2 \P_0.
+  \P_0 \O(w^\Delta,\hat{X})^2 \P_0.
 \end{align}
@@
 
-Here we construct the matrix $g^{(M)}$ that enables decomposing $M$-body operators $\O(w,\hat X)$ into terms $\O(w^\Delta,\hat X)$ that generate states of definite excitation energy $\Delta$ above the PS manifold, as in Eq.~\eqref{eq:excitation}.
+Here we construct the matrix $g^{(M)}$ that enables decomposing $M$-body operators $\O(w,\hat{X})$ into terms $\O(w^\Delta,\hat{X})$ that generate states of definite excitation energy $\Delta$ above the PS manifold, as in Eq.~\eqref{eq:excitation}.
 We work through the calculation of $g^{(1)}$ explicitly, and provide the result for $g^{(M)}$ from a generalized version of the same calculation.
-To this end, we consider the action of a single-body operator $\O(w,\hat X)=\sum_iw_i\hat X_i$ on an arbitrary PS state $\ket\psi\in\E_0$ and expand
-\begin{align}
-  \H_0 \O(w,\hat X) \ket\psi
-  = \f12 \sum_{i\ne j} \sum_k g_{ij} w_k \hat\Pi_{ij} \hat X_k \ket\psi,
+To this end, we consider the action of a single-body operator $\O(w,\hat{X})=\sum_iw_i\hat{X}_i$ on an arbitrary PS state $\ket\psi\in\E_0$ and expand
+\begin{align}
+  \H_0 \O(w,\hat{X}) \ket\psi
+  = \f12 \sum_{i\ne j} \sum_k g_{ij} w_k \hat\Pi_{ij} \hat{X}_k \ket\psi,
   \label{eq:single_body_diagnosis_start}
@@
 The sum in Eq.~\eqref{eq:single_body_diagnosis_start} has terms with $k\in\set{i,j}$ and terms with $k\notin\set{i,j}$.
-In the case of $k\notin\set{i,j}$, the permutation operator $\hat\Pi_{ij}$ commutes with $\hat X_k$ and annihilates on $\ket\psi$, and we can replace the sum
+In the case of $k\notin\set{i,j}$, the permutation operator $\hat\Pi_{ij}$ commutes with $\hat{X}_k$ and annihilates on $\ket\psi$, and we can replace the sum
 \begin{align}
@@
   \f12 \sum_{i\ne j} \sum_{k\notin\set{i,j}}
-  g_{ij} w_k \hat\Pi_{ij} \hat X_k \ket\psi
-  = E_0 \O(w,\hat X) \ket\psi
-  - \f12 \sum_{i\ne j} \sum_{k\in\set{i,j}} g_{ij} w_k \hat X_k \ket\psi,
+  g_{ij} w_k \hat\Pi_{ij} \hat{X}_k \ket\psi
+  = E_0 \O(w,\hat{X}) \ket\psi
+  - \f12 \sum_{i\ne j} \sum_{k\in\set{i,j}} g_{ij} w_k \hat{X}_k \ket\psi,
 \end{align}
@@
   \f12 \sum_{i\ne j} \sum_{k\notin\set{i,j}}
-  g_{ij} w_k \hat\Pi_{ij} \hat X_k \ket\psi
-  = E_0 \O(w,\hat X) \ket\psi - \sum_k g_k w_k \hat X_k \ket\psi.
+  g_{ij} w_k \hat\Pi_{ij} \hat{X}_k \ket\psi
+  = E_0 \O(w,\hat{X}) \ket\psi - \sum_k g_k w_k \hat{X}_k \ket\psi.
 \end{align}
@@
   \f12 \sum_{\substack{i\ne j\\k\in\set{i,j}}}
-  g_{ij} w_k \hat\Pi_{ij} \hat X_k \ket\psi
-  = \sum_{i,j} g_{ij} w_j \hat X_i \ket\psi.
+  g_{ij} w_k \hat\Pi_{ij} \hat{X}_k \ket\psi
+  = \sum_{i,j} g_{ij} w_j \hat{X}_i \ket\psi.
 \end{align}
@@
 \begin{align}
-  \H_0 \O(w,\hat X) \ket\psi
-  = E_0 \O(w,\hat X) \ket\psi
-  + \sum_k \sp{\sum_j g_{kj} w_j - g_k w_k} \hat X_k \ket\psi.
-\end{align}
-The action of the single-body perturbation $\O(w,\hat X)$ on a permutationally symmetric state therefore generates an eigenstate of $\H_0$ with interaction energy $E_0+\Delta$ if the vector $w=\sum_k w_k \ket{k}$ satisfies the eigenvalue equation
+  \H_0 \O(w,\hat{X}) \ket\psi
+  = E_0 \O(w,\hat{X}) \ket\psi
+  + \sum_k \sp{\sum_j g_{kj} w_j - g_k w_k} \hat{X}_k \ket\psi.
+\end{align}
+The action of the single-body perturbation $\O(w,\hat{X})$ on a permutationally symmetric state therefore generates an eigenstate of $\H_0$ with interaction energy $E_0+\Delta$ if the vector $w=\sum_k w_k \ket{k}$ satisfies the eigenvalue equation
 \begin{align}
@@
   \H_{\t{drive}}^{\t{full}}
-  = \sum_{j,\v v,\sigma} \Omega_{\v v\sigma}
-  \p{e^{-\i\kappa\v v\cdot\v\ell j} \s_{\v v\sigma j}^{(r)}
+  = \sum_{j,\bm{v},\sigma} \Omega_{\bm{v}\sigma}
+  \p{e^{-\i\kappa\bm{v}\cdot\bm\ell j} \s_{\bm{v}\sigma j}^{(r)}
     \otimes\op{\e}{\g}_j + \t{h.c.}}
-  + \Delta \hat N_\e,
-\end{align}
-where $\Omega_{\v v\sigma}$ is the amplitude of $\sigma$-polarized light propagating along axis $\v v$, with $\sigma=+1$ and $-1$ respectively for right and left circular polarizations; and $\s_{\v v\sigma j}$ is a spin-raising/lowering operator for atom $j$ along axis $\v v$, defined by appropriately rotating the single-atom spin operators
+  + \Delta \hat{N}_\e,
+\end{align}
+where $\Omega_{\bm{v}\sigma}$ is the amplitude of $\sigma$-polarized light propagating along axis $\bm{v}$, with $\sigma=+1$ and $-1$ respectively for right and left circular polarizations; and $\s_{\bm{v}\sigma j}$ is a spin-raising/lowering operator for atom $j$ along axis $\bm{v}$, defined by appropriately rotating the single-atom spin operators
 \begin{align}
   \s_\pm^{(r)}
-  \equiv -\sqrt{\f{n(n+1)(n-1)}{6}} \times \hat T_\pm^{(r)},
-  &&
-  \hat T_\pm^{(r)} \equiv \mp\sqrt{\f{2(s+r)+1}{2\ell+1}}
+  \equiv -\sqrt{\f{n(n+1)(n-1)}{6}} \times \hat{T}_\pm^{(r)},
+  &&
+  \hat{T}_\pm^{(r)} \equiv \mp\sqrt{\f{2(s+r)+1}{2\ell+1}}
   \sum_\mu \bk{s\mu;1,\pm1|s+r,\mu\pm1} \op{\mu\pm1}{\mu}.
 \end{align}
-Here $\bk{j_1m_1;j_2m_2|j_3m_3}$ is a Clebsch-Gordan coefficient, and we have normalized $\hat T_\pm^{(r)}$ such that $\tr\sp{\hat T_\pm^{(r)}{}^\dag \hat T_\pm^{(r)}}=1$.
-Still assuming real drive amplitudes, the corresponding effective drive Hamiltonian that replaces Eq.~\eqref{eq:H_3LD_single} in the far-detuned limit $\abs{\Delta}\gg\abs{\Omega_{\v v\sigma}}$ is then
+Here $\bk{j_1m_1;j_2m_2|j_3m_3}$ is a Clebsch-Gordan coefficient, and we have normalized $\hat{T}_\pm^{(r)}$ such that $\tr\sp{\hat{T}_\pm^{(r)}{}^\dag \hat{T}_\pm^{(r)}}=1$.
+Still assuming real drive amplitudes, the corresponding effective drive Hamiltonian that replaces Eq.~\eqref{eq:H_3LD_single} in the far-detuned limit $\abs{\Delta}\gg\abs{\Omega_{\bm{v}\sigma}}$ is then
 \begin{align}
@@
 \begin{align}
-  \H_\spin = -\f{u}{2N}\v\S\cdot\v\S + 2J\phi \sum_q \sin\p{q} \s_{\z,q}
+  \H_\spin = -\f{u}{2N}\bm\S\cdot\bm\S + 2J\phi \sum_q \sin\p{q} \s_{\z,q}
 \end{align}
@@
 \end{align}
-In the case of uniform SU($n$)-symmetric interactions of the form $\frac{g}{2}\v\S\cdot\v\S$ and a diagonal external field, we have
+In the case of uniform SU($n$)-symmetric interactions of the form $\frac{g}{2}\bm\S\cdot\bm\S$ and a diagonal external field, we have
 \begin{align}
@@
   + 2J\phi \sum_q \sin\p{q} \s_{\z,q},
+  \label{eq:H_spin_lax}
 \end{align}
 where $\S_{\mu\nu} = \sum_q \s_{\mu\nu q}$.
-The single-body operators that appear in this Hamiltonian have squared norms
+The Lax formalism \cite{yuzbashyan2005nonequilibrium, yuzbashyan2006dynamical, yuzbashyan2006relaxation, yuzbashyan2015quantum, smale2019observation} for analyzing a Hamiltonian of this form constructs a polynomial constant of motion, parameterized by a single (arbitrary) complex number.
+This polynomial has $N$ residues (where $N$ is the number of spins) corresponding to mutually commuting quantities whose appropirately weighted sum is equal to $\H_\spin$.
+When $n=2$, conservation of these residues provides sufficient dynamical constraints to make the spin system fully integrable.
+However, the size of Hilbert space grows with $n$, while the number of conserved quantities provided by the Lax analysis (namely, $N$) does not.
+In fact, a straightforward generalization of the Lax formalism to $n>2$ makes predictions that are inconsistent with the mean-field results in Figures \ref{fig:mean_mag_int} and \ref{fig:crit_fields_X} of the main text.
+We illustrate this claim with a direct calculation below, noting that this inconsistency is not a failure of the Lax formalism, but rather an indication that new theoretical tools are necessary to understand multilevel ($n>2$) spin models.
+
+The single-body operators that appear in Eq.~\eqref{eq:H_spin_lax} have squared norms
 \begin{align}
@@
   = \f1{12} (n+1) n (n-1)
-  \equiv \xi^2.
-\end{align}
-The Lax formulation (following Refs.~\cite{yuzbashyan2005nonequilibrium, yuzbashyan2006dynamical, yuzbashyan2006relaxation, yuzbashyan2015quantum, smale2019observation}) requires all single-body operators involved to have the same normalization, so we substitute $\s_{\tilde\z,q}\equiv \s_{\z,q}/\xi$ to expand
+  \equiv \xi^2,
+\end{align}
+whereas Lax formulation requires all single-body operators to have the same normalization.
+We therefore substitute $\tilde{s}_{\z,q}\equiv \s_{\z,q}/\xi$ to expand
 \begin{align}
@@
   = -\f1{2N} \sum_{\mu,\nu} \S_{\mu\nu} \S_{\nu\mu}
-  + \xi h \sum_q \sin\p{q} \s_{\tilde\z,q},
+  + \xi h \sum_q \sin\p{q} \tilde{s}_{\z,q},
   &&
@@
   \ell_\alpha\p{z}
-  = \f1N \sum_q \f{\s_{\alpha,q}}{z-\sin q}
-  + \delta_{\alpha,\tilde\z} \, \xi h,
-\end{align}
-where $\alpha$ indexes elements of a basis $\set{\s_\alpha}$ of self-adjoint generators of SU($n$), with normalization $\tr\p{\s_\alpha^2}=1$.
+  = \f1N \sum_q \f{\tilde{s}_{\alpha,q}}{z-\sin q}
+  + \delta_{\alpha,\z} \, \xi h,
+\end{align}
+where $\alpha$ indexes elements of a basis $\set{\tilde{s}_\alpha}$ of self-adjoint generators of SU($n$) that have normalization $\tr\p{\tilde{s}_\alpha^2}=1$, and $\delta_{\alpha,\z}=1$ if $\alpha=\z$ and $0$ otherwise.
 The squared magnitude $\vec\ell\p{z}^2=\sum_\alpha\ell_\alpha\p{z}^2$ is a constant of motion (for any $z$), and its residues provide $N$ mutually commuting quantities whose weighted sum recovers $\H_\spin$.
-When $n=2$, conservation of these residues provides sufficient dynamical constraints to make the spin system fully integrable.
-In this case, dynamical behavior is governed by the roots of $\vec\ell\p{z}^2$, and the presence (or absence) of complex roots marks distinct dynamical phases of $\H_\spin$.
-However, the size of Hilbert space grows with $n$, while the number of conserved quantities provided by the Lax analysis (namely, $N$) does not.
-When $n>2$, there is therefore no guarantee that the roots of $\vec\ell\p{z}^2$ will similarly govern dynamical behavior.
-In fact, a straightforward generalization of the Lax analysis to $n>2$ makes predictions that are inconsistent with the mean-field results in Figures \ref{fig:mean_mag_int_X}--\ref{fig:mean_mag_int_XXI} of the main text.
-We substantiate this claim with a direct calculation of the roots of $\vec\ell\p{z}^2$ below.
-
-Within the permutationally symmetric manifold, we can replace $\s_{\alpha,q}\to\bar s_\alpha\equiv\frac1N\sum_q \s_{\alpha,q}$ at the cost of $O(1/N)$ errors that vanish as $N\to\infty$, so taking this limit we find
+
+Within the permutationally symmetric manifold, we can replace $\tilde{s}_{\alpha,q}\to\bar{s}_\alpha\equiv\frac1N\sum_q \s_{\alpha,q}$ at the cost of $O(1/N)$ errors that vanish as $N\to\infty$, so taking this limit we find
 \begin{align}
   \ell_\alpha\p{z}
-  = \I\p{z} \bar s_\alpha
-  + \delta_{\alpha,\tilde\z} \, \xi h,
+  = \I\p{z} \bar{s}_\alpha
+  + \delta_{\alpha,\z} \, \xi h,
 \end{align}
@@
   = \sum_\alpha \ell_\alpha\p{z}^2
-  = \I\p{z}^2 \sum_{\alpha\ne\tilde\z} \bar s_\alpha^2
-  + \sp{\I\p{z} \bar s_{\tilde\z} + \xi h}^2,
-\end{align}
-where we can define the scalar $Q^2\equiv\sum_\alpha \bar s_\alpha^2$ to simplify
+  = \I\p{z}^2 \sum_{\alpha\ne\z} \bar{s}_\alpha^2
+  + \sp{\I\p{z} \bar{s}_\z + \xi h}^2,
+\end{align}
+where we can define the scalar $Q^2\equiv\sum_\alpha \bar{s}_\alpha^2$ to simplify
 \begin{align}
   \vec\ell\p{z}^2
-  = \I\p{z}^2 \p{Q^2 - \bar s_{\tilde\z}^2}
-  + \sp{\I\p{z} \bar s_{\tilde\z} + \xi h}^2
+  = \I\p{z}^2 \p{Q^2 - \bar{s}_\z^2}
+  + \sp{\I\p{z} \bar{s}_\z + \xi h}^2
   = \I\p{z}^2 Q^2 + \xi^2 h^2
-  + 2 \I\p{z} \xi h \bar s_{\tilde\z}.
-\end{align}
-For initial states with $\bk{\bar s_\z}=0$, we thus find that
+  + 2 \I\p{z} \xi h \bar{s}_\z.
+\end{align}
+For initial states with $\bk{\bar{s}_\z}=0$, we thus find that
 \begin{align}
@@
 \end{align}
-which is zero when\footnote{Strictly speaking, the zeros in Eq.~\eqref{eq:lax_zeros} occur at values of $z$ at which $\I\p{z}$ is undefined.
-We avoid this issue by analytically continuing $\I\p{z}^2$ to the interval $z\in(-1,1)$.}
+which is zero when\footnote{
+Strictly speaking, the zeros in Eq.~\eqref{eq:lax_zeros} occur at values of $z$ at which $\I\p{z}$ is undefined.
+We avoid this issue by analytically continuing $\I\p{z}^2$ to the interval $z\in(-1,1)$.
+}
 \begin{align}
@@
 \begin{align}
-  Q^2 = \sum_\alpha \bar s_\alpha^2
-  = \sum_{\mu,\nu} \bar s_{\mu\nu} \bar s_{\nu\mu} - \f1n
+  Q^2 = \sum_\alpha \bar{s}_\alpha^2
+  = \sum_{\mu,\nu} s_{\mu\nu} s_{\nu\mu} - \f1n
   = 1 - \f1n
@@
 This Lax analysis correctly predicts that $h_\crit=1$ when $n=2$, but otherwise predicts $h_\crit\sim n^{-3/2}$, which is inconsistent with the finding that $h_\crit\sim n^{-1/3}$ in the mean-field results of the main text (see Figure \ref{fig:crit_fields_X}).
-We emphasize that this inconsistency is not a failure of the Lax formalism, but rather an indication that new theoretical tools are necessary to understand multilevel spin models.
+This inconsistency is not a failure of the Lax formalism, but rather an indication that new theoretical tools are necessary to understand multilevel ($n>2$) spin models.
 

### 2022-10-04 sun_phases/spin_model_paper/figures/make_plots.py

@@ -41,249 +41,249 @@ def add_arrow(start, end, axis = None, *args, **kw_args):
     arrow = patches.FancyArrowPatch(start, end, *args, **kw_args)
     axis.add_patch(arrow)
 
-##########################################################################################
-# lattice diagram
-##########################################################################################
+# ##########################################################################################
+# # lattice diagram
+# ##########################################################################################
 
-figsize = (2,0.75)
-plt.figure(figsize = figsize)
-x_lim, y_lim = 1.6, 0.5
-x_vals = np.linspace(-x_lim, x_lim, 300)
-plt.plot(x_vals, -y_lim*np.cos(2*np.pi*x_vals), "k")
+# figsize = (2,0.75)
+# plt.figure(figsize = figsize)
+# x_lim, y_lim = 1.6, 0.5
+# x_vals = np.linspace(-x_lim, x_lim, 300)
+# plt.plot(x_vals, -y_lim*np.cos(2*np.pi*x_vals), "k")
 
-btm_y, top_y = -y_lim*0.4, y_lim*0.7
-plt.plot([-1], [btm_y], **marker_kw(dn_color))
-plt.plot([+0], [btm_y], **marker_kw(dn_color))
-add_arrow((-1,btm_y+0.08), (+0,btm_y+0.08),
-      connectionstyle="arc3,rad=-0.4", **single_kw)
-plt.annotate(r"$J$", (-0.5, btm_y), ha = "center", va = "center")
+# btm_y, top_y = -y_lim*0.4, y_lim*0.7
+# plt.plot([-1], [btm_y], **marker_kw(dn_color))
+# plt.plot([+0], [btm_y], **marker_kw(dn_color))
+# add_arrow((-1,btm_y+0.08), (+0,btm_y+0.08),
+#       connectionstyle="arc3,rad=-0.4", **single_kw)
+# plt.annotate(r"$J$", (-0.5, btm_y), ha = "center", va = "center")
 
-y_vals = np.linspace(top_y, btm_y, 30)
-x_vals = 1+0.06*np.sin(5*np.pi*(y_vals-top_y)/(top_y-btm_y))
-plt.plot(x_vals, y_vals, "r", linewidth = 1)
-plt.plot([1], [btm_y], **marker_kw(dn_color))
-plt.plot([1], [top_y], **marker_kw(up_color))
[…]
